### 2018-12-19 notes/graphtestingmasternotes.tex
added master notes

\documentclass{article}
\usepackage{amsmath, amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Var}[1]{\mathrm{Var}\left( #1 \right)}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\dotp}[2]{\langle #1 , #2 \rangle}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\Perp}{\perp \! \! \! \perp}

\newcommand{\Linv}{L^{\dagger}}
\newcommand{\tr}{\text{tr}}
\newcommand{\h}{\textbf{h}}
% \newcommand{\l}{\ell}
\newcommand{\x}{\textbf{x}}
\newcommand{\y}{\textbf{y}}
\newcommand{\Lx}{\mathcal{L}_X}
\newcommand{\Ly}{\mathcal{L}_Y}
\DeclareMathOperator*{\argmin}{argmin}

\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Rd}{\Reals^d}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}
\newtheorem{remark}{Remark}

\begin{document}

\title{Graph Testing Notes}
\author{Alden Green}
\date{\today}
\maketitle

\section{Setup}

\paragraph{Data model.}

We are given two distributions, $P$ and $Q$, with the ability to sample from either one. Our goal is to test the hypothesis $H_0: P = Q$ vs. the alternative $H_a: P \neq Q$.

Under the \textbf{binomial data model}, our sampling procedure is to draw i.i.d Rademacher labels $L_i \in \set{1, -1}$ for $i \in \set{1, \ldots, N}$, and then sample 
$Z_i \sim P$ if $L_i = 1$ and $Z_i \sim Q$ otherwise. Define $1_X$ to be the length-$N$ indicator vector for $L_i = 1$
\begin{equation*}
1_X[i] = 
\begin{cases}
1, L_i = 1\\
0 \text{ otherwise } 
\end{cases}
\end{equation*}
and similarly for $1_Y$
\begin{equation*}
1_Y[j] = 
\begin{cases}
1, L_i= -1 \\
0 \text{ otherwise } 
\end{cases}
\end{equation*}
and define $a = \frac{1_X}{N / 2} - \frac{1_Y}{N / 2}$.

Under the \textbf{fixed label data model} we use the same data generating process as above, except fix $\mathcal{L}_X = \set{1, \ldots, N/2}$ and $\mathcal{L}_Y = \set{N/2, \ldots, N}$. Say that $L_i = 1$ for $i \in \mathcal{L}_X$ and $L_i = -1$ for $i \in \mathcal{L}_Y$, and call $\set{ X_1, \ldots, X_{\abs{\Lx}} } = \set{Z_i: i \in \mathcal{L}_X}$ and likewise for $Y$.

\paragraph{Graph.}

Form an $N \times N$ Gram matrix $A$, where $A_{ij} = K(Z_i, Z_j)$ for \textbf{kernel function} $K$. Let $G = (V,E)$ with $V = \set{Z_1, \ldots, Z_n}$ and $E = \set{A_{ij}: 1 \leq i < j \leq n}$. Take $L = D - A$ to be the (unnormalized) \textbf{Laplacian matrix} of $A$ (where $D$ is the diagonal degree matrix with $D_{ii} = \sum_{j \in [n + m]} A_{ij}$). Denote by $B$ the $M \times N$ \textbf{incidence matrix} of $A$, where we denote the $i$th row of $B$ as $B_i$ and set $B_i$ to have entry $A_{ij}$ in position $i$, $-A_{ij}$ in position $j$, and $0$ everywhere else.

\paragraph{Resistance distances.}

There are many distances one can define over nodes in a graph. The \textbf{resistance distance between nodes $u$ and $v$}, $R_{uv}$, is defined as
\begin{equation*}
R_{uv} = (e_u - e_v)^T \Linv (e_u - e_v).
\end{equation*}

\paragraph{Test statistics.}

We begin by defining our \textbf{laplacian smooth} test statistic. 
\begin{equation*}
T_2 = \left(\max_{\theta: \norm{B\theta}_2 \leq 1} a^T \theta \right)^2 = a^T \Linv a. 
\end{equation*}

(Bhattacharya 2018) defines a general notion of 2-sample \textbf{graph-based test statistics}
\begin{equation*}
T_{\mathcal{G}} = \frac{1}{N^2}\sum_{i = 1}^{n} \sum_{j = n + 1}^{n + m} A_{ij}
\end{equation*}
Although he develops theory for this statistic in the context of $k$NN and minimum spanning tree graphs, we will at present consider it for the complete weighted similarity graph defined by $A$ above. Then, we can write 
\begin{equation*}
T_{\mathcal{G}} = a^T L a.
\end{equation*}

Finally, define $\mathcal{H}$ to be a \textbf{reproducing kernel Hilbert space} with $K$ the associated kernel. Let $\mathcal{F}$ be the unit ball of $\mathcal{H}$, and let the evaluation of $f \in \mathcal{F}$ at the sample points $Z_1, \ldots, Z_N$ be denoted by $\textbf{f} = f(Z_1, \ldots, Z_N)$. Then, the statistic $\text{MMD}_b$ of (Gretton 2012) can be written as
\begin{equation*}
T_{\mathcal{K}} = \sup_{f \in \mathcal{F}} a^T \textbf{f} = a^T K a. 
\end{equation*}

\paragraph{Distances between probability measures.}

We will need distances between probability measures for two different purposes. The first is that they are self-evidently useful in analyzing limiting distributions of statistics (in particular in this case, our test statistics).

For a function $f$, define its \textbf{Lipschitz norm} $\norm{f}_{L}$ to be
\begin{equation*}
\inf \set{K : \abs{f(x) - f(y)} \leq K \norm{x - y}}.
\end{equation*}
Define the \textbf{Wasserstein distance} between two measures $\mu$ and $\nu$ to be
\begin{equation*}
\mathcal{W}(\mu, \nu) := \sup \left \{ \abs{\int h \, d\mu - \int h \, d\nu}: h \text{ Lipschitz, with } \norm{h}_{L} \leq 1 \right\}.
\end{equation*}

If the measures $\mu$ and $\nu$ have corresponding cumulative distribution functions $F_{\mu}$ and $F_{\nu}$ then we can define the \textbf{Kolmogorov-Smirnov distance} to be
\begin{equation*}
\norm{F_\mu - F_\nu}_{\infty} := \sup_{t} \abs{F_{\mu}(t) - F_{\nu}(t)}.
\end{equation*}

The second reason we will use distances between probability measures is that they themselves make for good test statistics!

An \textbf{integral probability metric} (IPM) with respect to a function class $\F$ is defined
\begin{equation*}
\underset{f \in \F}{\sup} ~ \Expect{f(X)} - \Expect{f(Y)}
\end{equation*}
for $X \sim P$, $Y \sim Q$.

Hereafter, we will assume $P$ and $Q$ are absolutely continuous with respect to Lebesgue measure, with density functions $p$ and $q$, respectively. Denote the \textbf{mixture density} by $\mu = \frac{p + q}{2}$.

Denote the \textbf{gradient} of a function $f$ by $\nabla_x$. Then we can define the \textbf{Sobolev semi-norm} and \textbf{dot product}, $\norm{f}_{W_0^{1,2}(\X, \mu^2)}$ and $\dotp{f}{g}_{W_0^{1,2}(\X, \mu^2)}$, by
\begin{equation*}
\dotp{f}{g}_{W_0^{1,2}(\X, \mu)} = \int_{\X} \dotp{\nabla_x f(x)}{\nabla_x g(x)}_{\Rd} \mu^2(x), ~~ \norm{f}_{W_0^{1,2}(\X, \mu)} = \sqrt{\int_{\X} \norm{\nabla_x f(x)}^2 \mu^2(x) dx }
\end{equation*}

Let the \textbf{Sobolev space}, $W^{1,2}(\mathcal{X}, \mu^2)$, be 
\begin{equation*}
W^{1,2}(\mathcal{X},\mu^2) = \set{f: \X \to \Reals, \int_{\X} \norm{\nabla_x f(x)}^2 \mu^2(x) dx < \infty}.
\end{equation*}
and denote by $W_0^{1,2}(\mathcal{X}, \mu^2)$ the restriction of $W^{1,2}(\mathcal{X}, \mu^2)$ to functions which vanish at the boundary of $\X$. Note that $\norm{f}_{W_0^{1,2}(\X, \mu^2)}$ defines a semi-norm over $W_0^{1,2}(\X, \mu^2)$. Finally, let $B_W(\X, \mu^2)$ be the \textbf{unit ball} of $W_0^{1,2}(\X, \mu^2)$, meaning
\begin{equation*}
B_W(\X, \mu^2) = \set{f \in W_0^{1,2}(\X, \mu^2): \norm{f}_{W_0^{1,2}(\X, \mu^2)} \leq 1}
\end{equation*}

Now we can define the \textbf{Sobolev IPM}, $\mathcal{S}_{\mu^2}(P,Q)$ It is simply an IPM where the function class is the Sobolev unit ball with respect to $\mu^2$.
\begin{equation*}
\mathcal{S}_{\mu^2}(P,Q) \overset{\mathrm{def}}{=} \sup_{f \in B_W} \biggl\{ \Expect{f(X)} - \Expect{f(Y)}\biggr\}
\end{equation*}

\paragraph{Holder functions.}
We will show that the Laplacian constraint $\norm{B \theta}_2 \leq 1$ is very similar to the constraint $f_{\theta} \in B_W(X,\mu^2)$ for the right choice of $K$, over all Holder functions.

For mapping $f: \Rd \to \Reals$ and $\beta$ a positive integer, we say $f$ is a \textbf{$\beta$-Holder function} if there exists $C > 0$ such that for all $x, y \in \X$
\begin{equation*}
\abs{f^{(\beta - 1)}(x) - f^{(\beta - 1)}(y)} \leq K \norm{x - y}
\end{equation*}
Roughly speaking, this means the functions have bounded $\beta$ partial derivatives.

\section{Conjectures}

Conjectures \ref{conj: spectral_measure_conv} and \ref{conj: diagonal_entries} will be needed for Theorem \ref{thm: asymptotic_null_distribution}.

\begin{conjecture}
	\label{conj: spectral_measure_conv}
	There exists a sequence of scaling factors $(\rho_n)_{n = 1}^{\infty}$ such that the \textbf{spectral measure} $\mu_n$ of $\rho_n \Linv$ converges weakly in probability
	\begin{equation*}
	\mu_n(\rho_n \Linv) \overset{\ast}{\rightharpoonup} \nu_{\infty}.
	\end{equation*}

	where $V \sim \nu_{\infty}$ and $V_n \sim \mu_n$ are bounded almost surely for all $n$ by some constant $C$. 
\end{conjecture}

\begin{conjecture}
	\label{conj: diagonal_entries}
	For all $\epsilon > 0$, there exists $N$ such that
	\begin{equation*}
	\Prob{\underset{i \in [n]}{\max} \, \frac{1}{n} \left(\{ \rho_n \Linv\}^2 \right)_{ii} \leq \epsilon} \geq 1 - \epsilon
	\end{equation*}
	for all $n \geq N$.
\end{conjecture}

\section{DESIRED RESULTS}

\begin{theorem}
	\label{thm:convprob_of_T2}
	For bandwidth parameter $h >0$ and decreasing function $k(\cdot,\cdot)$, write
	\begin{equation*}
	K(Z_i,Z_j) = \frac{1}{h^m} k(\norm{Z_i - Z_j}^2 / h^2).
	\end{equation*}

	For Sobolev IPM $\mathcal{S}_{\mu^2}(P,Q)$ as defined above,
	\begin{equation*}
	\sqrt{T_2} \convprob \mathcal{S}_{\mu^2}(P,Q)
	\end{equation*}
\end{theorem}

\begin{proof}[Proof attempt of Proposition \ref{thm:convprob_of_T2}]
	Recall that, for incidence matrix $B$, 
	\begin{equation*}
	\sqrt{T_2} = \left(\max_{\theta: \norm{B\theta}_2 \leq 1} a^T \theta \right).
	\end{equation*}

	We expand $\abs{\sqrt{T_2} - \mathcal{S}_{\mu^2}(P,Q)}$,
	\begin{align}
	\label{eqn: bias_variance_expansion}
	\abs{\sqrt{T_2} - \mathcal{S}_{\mu^2}(P,Q)} & \leq \abs{\max_{\theta: \norm{B\theta}_2 \leq 1} \bigl\{a^T \theta \bigr\} - \sup_{f \in B_W(\X, \mu^2)} \left\{ \mathbb{P}_n(f) - \mathbb{Q}_n(f) \right\} } \nonumber \\
	& + \abs{\sup_{f \in B_W(\X, \mu^2)} \left\{ \mathbb{P}_n(f) - \mathbb{Q}_n(f) \right\} - \sup_{f \in B_W(\X, \mu^2)} \left\{ \mathbb{P}(f) - \mathbb{Q}(f) \right\} }
	\end{align}

	\textcolor{red}{(The following statement would hold only if Proposition \ref{prop: sobolev_donsker} held over $B_W(\X, \mu^2)$, rather than over $B_W([0,1], \lambda)$ for $\lambda$ Lebesgue measure.)}

	By Proposition \ref{prop: sobolev_donsker}, the second term in the summand on the right hand side of (\ref{eqn: bias_variance_expansion}) is $o_P(1)$.

	\textcolor{red}{(The following statement would hold only if Proposition \ref{prop: convprob_regularization_functional} were uniform over $B_W(\X, \mu^2)$ rather than over the class of $\alpha$-Holder functions $\F_{\alpha}$)}

	Then, Proposition \ref{prop: convprob_regularization_functional} implies that for any $\epsilon > 0$, there exists $N$ such that for $n \geq N$, 
	\begin{equation*}
	\sup_{f \in B_W(\X, \mu^2)} \left\{ \mathbb{P}_n(f) - \mathbb{Q}_n(f) \right\} - \max_{\theta: \norm{B\theta}_2 \leq 1} \bigl\{a^T \theta \bigr\} \leq \epsilon
	\end{equation*}
	with high probability.

	To complete the proof, we will have to show that for any $\epsilon > 0$, there exists $N$ such that for $n \geq N$, 
	\begin{equation*}
	\max_{\theta: \norm{B\theta}_2 \leq 1} \bigl\{a^T \theta \bigr\} - \sup_{f \in B_W(\X, \mu^2)} \left\{ \mathbb{P}_n(f) - \mathbb{Q}_n(f) \right\} \leq \epsilon
	\end{equation*}
	with high probability.
\end{proof}

\section{Results}

\paragraph{Expectation of two-sample test statistics.}

The expectation of the above statistics is potentially a good way to understand their large sample behavior, as quadratic forms often satisfy laws of large numbers assuming the matrices are well-conditioned.

\begin{proposition}
	\label{prop: expectation_of_TG}
	Draw $Z$ and $a$ under the binomial data model, and assume both $P$ and $Q$ are absolutely continuous with respect to Lebesgue measure over Euclidean space $\Reals^d$. Write $h_0(x) = p(x) - q(x)$ with empirical analogue $\mathbf{h_0} = (h_0(Z_1), \ldots, h_0(Z_n) )$. Then
	\begin{align}
	\label{eqn: expectation_of_TG}
	\Expect{T_{\mathcal{G}}} = & \int \int K(\norm{\x - \y}) \left[p(\x) + q(\x)\right] \left[p(\y) + q(\y)\right] d\x d\y \nonumber \\
	- & \frac{N}{N - 1} \int \int K(\norm{\x - \y}) \left[h_0(\x)\right]^2 \frac{p(\y) + q(\y)}{p(\x) + q(\x)} d\x d\y \nonumber \\
	+ & \frac{N}{N - 1} \int \int K(\norm{\x - \y}) \left[h_0(\x) - h_0(\y)\right]^2 \left[p(\y) + q(\y)\right] \left[p(\x) + q(\x)\right] d\x d\y.
	\end{align}
\end{proposition}

Note that even under the null hypothesis, where $h_0 = 0$,
\begin{equation*}
\Expect{T_{\mathcal{G}}} = \int \int K(\norm{\x - \y}) \left[p(\x) + q(\x)\right] \left[p(\y) + q(\y)\right] d\x d\y
\end{equation*}
which is not distribution-free, unlike in the case of the $k$NN graph.

Another interesting consequence of Proposition \ref{prop: expectation_of_TG} comes when we take $K(\x,\y) = K(\frac{\norm{\x - \y}}{t})$ and let $t \to 0$.

\begin{proposition}
	\label{prop: expectation_of_TG_with_kde}
	If $p$ and $q$ are Lipschitz continuous functions with bounded hessians, and $K$ is a continuous function on $R^+$ such that $x^{2 + d} K(x) \in L_2$, then under the same setup as in Proposition \ref{prop: expectation_of_TG},
	\begin{equation}
	\label{eqn: expectation_of_TG_with_kde}
	\frac{N - 1}{N}\underset{t \to 0}{\lim} \frac{1}{t^d} \Expect{T_{\mathcal{G}}} = \int  h_0(\x)^2 d\x + \int (p(\x) + q(\x))^2 d\x
	\end{equation}
\end{proposition}

We turn now to the expectation of the Laplacian smooth statistic.
\begin{proposition}
	\label{prop: expectation_of_T2}
	Under the fixed label data model
	\begin{equation}
	\label{eqn: expectation_of_T2}
	\Expect{a^T \Linv a} = \Expect{R_{X_1 Y_1}} - \frac{N - 1}{2N} \Expect{R_{X_1,X_2}} - \frac{N - 1}{2N} \Expect{R_{Y_1, Y_2}}
	\end{equation}
\end{proposition}

Note that, although the resistance distances are between only two nodes, in each case the expectation is over the entire (random) graph $G$.

\paragraph{Asymptotic null distribution for $T_2$.}

We can compute an asymptotic null distribution for $T_2$, although its formulation depends on the eigenvalues of the matrix $\Linv$ which themselves are not obvious.

\begin{theorem}
	\label{thm: asymptotic_null_distribution}
	Denote the scaled version of the Laplacian smooth test statistic
	\begin{equation*}
	W_n = \sqrt{\frac{N^4}{32 \cdot \tr((\Linv)^2)}}\Bigl(T_2^2 - \frac{\tr(\Linv)}{4N^2} \Bigr).
	\end{equation*}

	If Conjectures \ref{conj: spectral_measure_conv} and \ref{conj: diagonal_entries} hold,
	\begin{equation*}
	\lim_{n \to \infty} \sup_{t} \abs{\Prob{W_n \leq t} - \Phi(t)} = 0.
	\end{equation*}
\end{theorem}

To prove Theorem \ref{thm: asymptotic_null_distribution}, we will need the following calculations of moments under $H_0$.

\begin{lemma}
	\label{lem: cond_expectation_of_T2}
	Under $H_0$, the conditional expectation $\Expect{T_2 \vert Z} = \frac{\tr(\Linv)}{N^2}$. 
\end{lemma}

\begin{lemma}
	\label{lem: cond_var_of_T2}
	Under $H_0$, the conditional variance $\Var{T_2 \vert Z} = \frac{32 \tr[(\Linv)^2]}{N^4}$. 
\end{lemma}

\section{Supplemental Results}

\paragraph{Empirical process over Sobolev classes.}

The following theorem is a stand-in; it handles only functions with domain on the unit interval, and is stated specifically with respect to Lebesgue measure.
\begin{proposition}
	\label{prop: sobolev_donsker}
	Let $\F$ be the set of all absolutely continuous functions $f: [0,1] \to \Reals$ such that $\norm{f}_{\infty} \leq 1$ such that $\int (f'(x))^2 dx \leq 1$. Then, there exists a constant $K$ such that for every $\epsilon > 0$,
	\begin{equation*}
	\log N_{[]}(\epsilon, \mathcal{F}, \norm{\cdot}_{\infty}) \leq K\left(\frac{1}{\epsilon}\right).
	\end{equation*}

	Thus, the class $\mathcal{F}$ is $P$-Donsker (and $P$-Glivenko-Cantelli) for all $P$.
\end{proposition}

\paragraph{Regularization functionals.}

When taking the supremum over functions which satisfy $\norm{B \theta}_2 \leq 1$, we will argue that this constraint is well-behaved in the limit, i.e. that it converges to the \textbf{regularization functional} $\norm{\cdot}_{W_0^{1,2}(\X, \mu^2)}$. Proposition \ref{prop: convprob_regularization_functional} makes this convergence uniform over the set of $3$-Holder functions (essentially functions with bounded $3rd$ derivative). Proposition \ref{prop: bousquet04} makes this convergence only pointwise, but merely requires that $f$ have bounded 2nd derivative.

\begin{proposition}
	\label{prop: convprob_regularization_functional}
	Let $\mathcal{F}_{\alpha}$ be a unit ball in the space of $\alpha$-Holder functions, and define $k(\cdot,\cdot)$ as in Theorem \ref{thm:convprob_of_T2}. For function $f \in \F_{\alpha}$, denote $f$ evaluated on the data, $\mathbf{f} = (f(Z_1), \ldots, f(Z_N))$. Then, there exists a constant $c$ depending only on $k$ such that for $\alpha \geq 3$ and a sequence $(h_n) \to 0$ such that
	\begin{equation*}
	\sup_{f \in \mathcal{F}_{\alpha}} \abs{ \norm{B \mathbf{f}_2} - \norm{f}_{W_0^{1,2}(\X, \mu^2)} } \convprob 0
	\end{equation*}
\end{proposition}

\begin{proposition}[Bousquet 04]
	\label{prop: bousquet04}
	If $p$ and $q$ are Lipschitz continuous functions with bounded hessians, and $K$ is a continuous function on $R^+$ such that $x^{2 + d} K(x) \in L_2$, then 
	\begin{align}
	\label{eqn: bousquet04}
	& \underset{t \to 0}{\lim} \frac{d}{Ct^{d+2}} \int  K(\norm{\x - \y}/ t)(h_0(\x) - h_0(\y))^2 (p(\x) + q(\x))(p(\y) + q(\y)) d\x d \y \nonumber \\
	& = \int \norm{\nabla h_0(\x)}^2 (p(\x) + q(\x))^2 d\x
	\end{align}
\end{proposition}

\begin{lemma}[von Luxburg 12]
	\label{lem: vonluxburg12}
	Assume $P$ and $Q$ are absolutely continuous with respect to Lebesgue measure on Euclidean space $\Reals^d$, with density functions $p$ and $q$, respectively. Let $K(x,y) = \frac{1}{(2\pi \sigma^2)}^{d/2} \exp{-\frac{\norm{x - y}^2}{2 \sigma^2}}$.

	Under some regularity assumptions on $p$ and $q$, if $n \to \infty$, $\sigma \to 0$ , and $n\sigma^{d+2}/ \log(n) \to \infty$, then
	\begin{align*}
	n R_{XY} \to \frac{2}{p(X) + q(X)} + \frac{2}{p(Y) + q(Y)} \textit{ almost surely }
	\end{align*}
	with equivalent statements holding for $X_1, X_2$ and $Y_1, Y_2$.  
\end{lemma}

\paragraph{Central limit theorem for quadratic forms.}

\begin{theorem}[Chatterjee 08]
	\label{thm: clt_quadratic_forms}
	Let $a = (a_1, \ldots, a_n)$ be i.i.d random variables with with $\Prob{X_i = 1} = \Prob{X_i = -1} = 1/2$. For some fixed real valued symmetric matrix $M = (M_{ij})_{1 \leq i, j \leq n}$, define
	\begin{equation*}
	W = a^T M a.
	\end{equation*}
	with $\mu$ denoting the law of $(W - EW) / \sqrt{\Var{W}}$.

	Then, letting $\mathcal{G}$ be the standard Gaussian measure 
	\begin{equation}
	\label{eqn: wass_distance_to_normal}
	\mathcal{W}(\mu, \mathcal{G} ) \leq \left(\frac{\tr(M^4)}{\tr(M^2)^2} \right)^{1/2} + \left( \frac{5 \, \underset{i}{\max} \, (M_{ii})^2}{\tr(M^2)} \right)^{1/2}.
	\end{equation}
\end{theorem}

\paragraph{Translating from Wasserstein to Kolmogorov distance.}

\begin{lemma}[Wasserstein to Kolmogorov distance]
	\label{lem: wass_to_ks_distance}
	For any probability measures $\mu$, $\nu$ with corresponding cdfs $F_{\mu}$ and $F_{\nu}$ and any $\epsilon' > 0$, there exists some $\epsilon > 0$ such that
	\begin{equation*}
	\mathcal{W}(\mu, \nu ) < \epsilon \implies \sup_{t} \abs{F_{\mu}(t) - F_{\nu}(t)} \leq \epsilon'.
	\end{equation*}
\end{lemma}

\section{Proofs}

\begin{proof}[Proof of Proposition \ref{prop: expectation_of_TG}]
	Throughout, we will use the fact that $a_i | Z_i \sim \text{Rademacher}(\frac{p(Z_i)}{p(Z_i) + q(Z_i)})$, which is easily seen by an application of Bayes rule.

	Begin by rewriting
	\begin{equation*}
	a^T L a = (\mathbf{h_0} + a - \mathbf{h_0})^T L (\mathbf{h_0} + a - \mathbf{h_0}) := (\mathbf{h_0} + \mathbf{\epsilon})^T L (\mathbf{h_0} + \mathbf{\epsilon}).
	\end{equation*}
	Expanding the quadratic form yields
	\begin{equation*}
	a^T L a = \mathbf{h_0}^T L \mathbf{h_0} + \mathbf{\epsilon}^T L \mathbf{\epsilon} + 2 \mathbf{h_0}^T L \mathbf{\epsilon}.
	\end{equation*}

	Going from back to front, we have that the first term has expectation $0$, because
	\begin{equation*}
	\Expect{L_{ij} h_0(Z_i) \epsilon_{j}} = \Expect{L_{ij} h_0(Z_i) \Expect{\epsilon_{j} | Z} } = \Expect{L_{ij} h_0(Z_i) 0} = 0.
	\end{equation*}

	For the middle term, only the diagonal terms have non-zero expectation.
	\begin{align*}
	\Expect{ \mathbf{\epsilon}^T L \mathbf{\epsilon} } & = \sum_{i,j = 1}^{N} \Expect{L_{ij} \Expect{\epsilon_{j} \epsilon_i | Z}} \\
	& \overset{(i)}{=} \sum_{1 \leq i < j \leq N} \Expect{L_{ij} \Expect{\epsilon_{j} | Z} \Expect{\epsilon_i | Z}} + \sum_{i = 1}^n \Expect{L_{ii}^2 \Expect{\epsilon_i^2 | Z}} \\
	& = \sum_{i = 1}^N \Expect{L_{ii}^2 \Expect{\epsilon_i^2 | Z}}.
	\end{align*}
	where $(i)$ follows from the conditional independence relation $a_i \Perp a_j | Z$.

	Then 
	\begin{equation*}
	\Expect{\epsilon_i^2 | Z} = \Expect{(a(Z_i) - h_0(Z_i))^2 | Z} = \Var{a(Z_i) | Z_i} = \frac{4}{N^2} \left(\frac{4 p(Z_i) q(Z_i)}{(p(Z_i) + q(Z_i))^2}\right)
	\end{equation*}
	and plugging this in, we have
	\begin{align*}
	\Expect{ \mathbf{\epsilon}^T L \mathbf{\epsilon} } & = \frac{16}{N^2}\sum_{i = 1}^{N} \Expect{L_{ii} \left(\frac{p(Z_i) q(Z_i)}{(p(Z_i) + q(Z_i))^2}\right)} \\
	& = \frac{16}{N^2} \sum_{i = 1}^{N} \Expect{ \sum_{j \neq i}K(\norm{Z_i - Z_j}) \left(\frac{p(Z_i) q(Z_i)}{(p(Z_i) + q(Z_i))^2}\right)} \\
	& = \frac{4 N}{N - 1} \int \int K(\norm{\x - \y}) \left[p(\x) q(\x)\right] \frac{p(\y) + q(\y)}{p(\x) + q(\x)} d\x d\y \\
	\end{align*}

	Using the relation $\frac{(a + b)^2 - (a - b)^2}{4} = ab$ yields the 1st and 2nd integrals of (\ref{eqn: expectation_of_TG}). The 3rd integral is exactly $\Expect{h_0^T L h_0}$.

\end{proof}

\begin{proof}[Proof of Proposition \ref{prop: expectation_of_TG_with_kde}]
	Write $\mathbf{h} = \frac{\x - \y}{t}$. Via Taylor expansion, we can write
	\begin{align*}
	& \int K\left(\frac{\norm{\x - \y}}{t}\right) (p(\y) + q(\y) ) d\y \\
	& \overset{(i)}{=} \int K(\norm{\mathbf{h}}) (p(\x) + q(\x) + \mathcal{O}(t \norm{h})) t^d d \mathbf{h} \\
	& \overset{(ii)}{=} (p(\x) + q(\x)) + \mathcal{O}(t^{d + 1})
	\end{align*}
	where $(i)$ follows from the Lipschitz continuity of $p$ and $q$, and $(ii)$ follows from the integrability condition on $K$.

	Applying this to the 2nd and 3rd integrals of (\ref{eqn: expectation_of_TG}) yields the two integrals of (\ref{eqn: expectation_of_TG_with_kde}). The 3rd integral is $\mathcal{O}(t^{d + 1})$ by Lemma \ref{prop: bousquet04}.
\end{proof}

\begin{proof}
	First, we rewrite $T_2$, using the fact that $a = \frac{2}{N} \left(\sum_{i \in \Lx} e_i - \sum_{i \in \Ly} e_i \right)$. 
	\begin{align*}
	a^T \Linv a & = \frac{4}{N^2} \left(\sum_{i, j \in \Lx }e_i \Linv e_j + \sum_{i, j \in \Ly }e_i \Linv e_j - 2\sum_{i \in \Lx, j \in \Ly} e_i \Linv e_j \right)\\
	\end{align*}
	Via this expression, we see that in the above summations
	\begin{itemize}
		\item For $i = j$, $e_i^T \Linv e_i$ appears exactly once.
		\item For $i \neq  j$ and $i, j \in \Lx$ or $i,j \in \Ly$, $e_i^T \Linv e_j$ appears exactly twice.
		\item For $i \in \Lx$, $j \in \Ly$, $- e_i^T \Linv e_j$ appears exactly twice.
	\end{itemize}

	Now, consider the expression
	\begin{equation*}
	 \sum_{u \in \Lx, v \in \Ly} R_{uv} - \sum_{u < v \in \Ly }R_{uv} - \sum_{u < v \in \Lx }R_{uv}.
	\end{equation*}
	Going from bottom to top, we have
	\begin{itemize}
		\item When $i \in \Lx$ and $j \in \Ly$, $R_{ij}$ will contribute $-2 e_i \Linv e_j$. No other $R_{uv}$ will contribute anything to this term.
		\item When $i < j \in \Lx$ or $i < j \in \Ly$, the term $R_{ij}$ in the 2nd or 3rd sum will appear exactly once and will contribute $2 e_i \Linv e_j$. No other $R_{uv}$ will contribute anything to this term.
		\item When $i = j \in \Lx$, $- R_{ik}$ will contribute $ - e_i \Linv e_i$ for each $k \neq i \in \Lx$, and will contribute $ e_i \Linv e_i$ for each $k \in \Ly$. The total contribution will be $ \left(\abs{\Ly} - \abs{\Lx} + 1\right) (e_i \Linv e_i) = e_i \Linv e_i$. The same reasoning holds for $i = j \in \Ly$. 
	\end{itemize}

	All contributions from all $R_{uv}$ can be put into one of the three proceeding categories. Therefore,
	\begin{equation*}
	a^T \Linv a = \frac{4}{N^2} \left(\sum_{u \in \Lx, v \in \Ly} R_{uv} - \sum_{u < v \in \Ly }R_{uv} - \sum_{u < v \in \Lx }R_{uv}\right)
	\end{equation*}

	(\ref{eqn: expectation_of_T2}) follows from taking expectation and noting that $X_i$ and $X_j$ are identically distributed for all $i$ and $j$.

\end{proof}

\begin{proof}[Proof of Theorem \ref{thm: asymptotic_null_distribution}]
	We will proceed by
	\begin{enumerate}
		\item Conditioning on the high-probability outcome that the Laplacian converges to a limiting object in the right sense.
		\item Showing that, under such convergence of the Laplacian, both terms in Theorem \ref{thm: clt_quadratic_forms} grow small with $n$.
		\item Converting from Wasserstein distance to Kolmogorov distance. 
	\end{enumerate}

	\paragraph{Step 1.}
	Fix $\epsilon > 0$. Throughout, let $P_Z$ denote the distribution of $Z$, and likewise $P_a$ denote the distribution of $a$.

	For  $V_n \sim \nu_n(\rho_n \Linv)$, and $V \sim \nu_{\infty}$ let
	\begin{equation*}
	A_n = \biggl \{z \in \mathbb{R}^n: \abs{ E V_n^p - E V^p } \leq \epsilon \text{ for } p = 1, 2, 4 \biggr \} \bigcup \set{ z \in \mathbb{R}^n: \underset{i \in [n]}{\max} \, \frac{1}{n} \left(\{ \rho_n \Linv\}^2 \right)_{ii} \leq \epsilon}.
	\end{equation*}

	It is not hard to see that our Conjectures \ref{conj: spectral_measure_conv} and \ref{conj: diagonal_entries} imply $A_n$ will eventually have high probability.
	\begin{align}
	\label{eqn: good_event_probability}
	\Prob{A_n} & \geq \Prob{\biggl \{z \in \mathbb{R}^n: \abs{ E V_n^p - E V^p } \leq \epsilon \biggr \}} + \Prob{\set{ z \in \mathbb{R}^n: \underset{i \in [n]}{\max} \, \frac{1}{n} \left(\{ \rho_n \Linv\}^2 \right)_{ii} \leq \epsilon}} \nonumber \\
	& \overset{(i)}{\geq} 1 - 2 \epsilon \text{ for all } n \geq N.
	\end{align}
	where $(i)$ follows from Conjecture \ref{conj: diagonal_entries} (for the second term), and Conjecture \ref{conj: spectral_measure_conv} (for the first term).

	Writing $W_n := W_n(z,a)$ to emphasize that it is a function of $z$ and $a$, we have by Tonelli's theorem that
	\begin{align}
	\label{eqn: conditional_ks_distance}
	\sup_{t} \abs{\Prob{W_n \leq t} - \Phi(t)} & \overset{(i)}{=} \sup_{t} \abs{ \int_{\Reals^N} \left( \int_{\{-1,1\}^N} 1(W_n(z,a) \leq t) dP_a\right) dP_z - \Phi(t) } \nonumber \\
	& = \sup_{t} \abs{ \int_{\Reals^N} \left( \int_{\{-1,1\}^N} 1(W_n(z,a) \leq t) dP_a\right) - \Phi(t)dP_z } \nonumber \\
	& \leq \int_{\Reals^N} \sup_{t} \abs{\left( \int_{\{-1,1\}^N} 1(W_n(z,a) \leq t)   dP_a\right) - \Phi(t) } dP_z \nonumber \\
	& \overset{(ii)}{\leq} \int_{A_n} \sup_{t} \abs{\left( \int_{\{-1,1\}^N} 1(W_n(z,a) \leq t) dP_a\right) - \Phi(t)} dP_z + 2 \epsilon
	\end{align}
	where $(i)$ follows from Tonelli's theorem and $(ii)$ from (\ref{eqn: good_event_probability}).

	\paragraph{Step 2.}

	Denote as
	\begin{equation*}
	F_{a|z}(z, t) := \left( \int_{\{-1,1\}^N} 1(W_n(z,a) \leq t) dP_a\right)
	\end{equation*}
	and note that for any $z$ this defines a measure over the Borel $\sigma$-algebra $\mathcal{B}(\Reals)$, which we will call $\mu_{a|Z}(z)$.

	We wish to upper bound $\mathcal{W}(\mu_{a|Z}(z), \mathcal{G})$. To do so, we will compute upper bounds for each present in (\ref{eqn: wass_distance_to_normal}). For the first term, we have
	\begin{align*}
	\frac{\tr(\{\Linv \}^4)}{\tr(\{ \Linv \}^2)^2} & = \frac{1}{n} \frac{\frac{1}{n} \tr(\rho_n^4 \{\Linv \}^4)}{ \frac{1}{n^2} \rho_n^4 \tr(\{ \Linv \}^2)^2} \\
	& \leq \frac{1}{n} \frac{\Expect{V^4} + \epsilon}{\Expect{V^2}^2 - \epsilon}.
	\end{align*}
	For the second term, we have
	\begin{align*}
	\frac{ \underset{i}{\max}(\{\Linv\}^2)_{ii} }{ \tr(\{\Linv\}^2) } & = \frac{ \frac{\rho_n^2}{n} (\{\Linv\}^2)_{ii}}{ \frac{\rho_n^2}{n} \tr(\{\Linv\}^2) } \\
	& \leq \frac{ \epsilon }{ \Expect{V^2} - \epsilon }.
	\end{align*}
	By Theorem \ref{thm: clt_quadratic_forms} we therefore have
	\begin{equation}
	\label{eqn: wasserstein_conditional_bound}
	\mathcal{W}(\mu_{a|Z}(z), \mathcal{G} ) \leq \frac{1}{n} \frac{\Expect{V^4} + \epsilon}{\Expect{V^2}^2 - \epsilon} + \left(\frac{ \epsilon }{ \Expect{V^2} - \epsilon } \right)^{1/2}.
	\end{equation}

	\paragraph{Step 3.}
	Note that the right hand side of (\ref{eqn: wasserstein_conditional_bound}) converges to 0 with $\epsilon$. Therefore, for any $\epsilon$ sufficiently small, by (\ref{eqn: wasserstein_conditional_bound}) and Lemma \ref{lem: wass_to_ks_distance} we have
	\begin{equation*}
	\norm{F_{Z | a} - \Phi}_{\infty} \leq \epsilon'.
	\end{equation*}
	Combined with (\ref{eqn: conditional_ks_distance}) we have
	\begin{equation*}
	\sup_{t} \abs{\Prob(W_n \leq t) - \Phi(t)} \leq 2 \epsilon + \epsilon'.
	\end{equation*}
	for all $n \geq n_0$.

\end{proof}

\begin{proof}[Proof of Lemma \ref{lem: cond_expectation_of_T2}]
	\begin{align}
	\Expect{T^2 \vert Z} & = \Expect{a^T \Linv a \vert Z} \nonumber \\
	& \overset{(i)}{=} \sum_{i = 1}^{N} \sum_{j = 1}^{N} \Expect{a_i a_j} \Linv_{ij} \nonumber \\
	& =  \sum_{i = 1}^{N} \frac{1}{4N^2} \Linv_{ii} \nonumber \\
	& = \frac{\tr(\Linv)}{4N^2}.
	\end{align}
	where $(i)$ comes from the independence of $Z$ and $a$ under $H_0$.
\end{proof}

\begin{proof}[Proof of Lemma \ref{lem: cond_var_of_T2}]
	First, we re-arrange $T_2$. 
	\begin{align*}
	T_2 & = \sum_{i = 1}^{N} \sum_{j = 1}^{N} a_i a_j \Linv_{ij} \\
	& = 2 \sum_{i \leq j} a_i a_j \Linv_{ij} - \frac{4}{N^2}\sum_{i = 1}^{N} \Linv_{ii}.
	\end{align*}
	Therefore, for $R_i \overset{i.i.d}{\sim} \text{Rademacher}(1/2)$,
	\begin{align*}
	\Var{T_2 \vert Z} & = 4 \Var{\sum_{i \leq j} a_i a_j \Linv_{ij} | Z} \\
	& = \frac{64}{N^4} \Var{\sum_{i \leq j} R_i R_j \Linv_{ij} \vert Z } \\
	& = \frac{32}{N^4} \tr[(\Linv)^2].
	\end{align*}
\end{proof}

\end{document}

### 2019-01-07 notes/weekly_notes/12.31.18.NotesForWeek.tex
added winter break notes

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\makeatletter
\newcommand{\leqnomode}{\tagsleft@true}
\newcommand{\reqnomode}{\tagsleft@false}
\makeatother

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Var}[1]{\mathrm{Var}\left( #1 \right)}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\dotp}[2]{\langle #1 , #2 \rangle}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\defeq}{\overset{\mathrm{def}}{=}}
\newcommand{\convweak}{\overset{w}{\rightharpoonup}}
\newcommand{\dive}{\mathrm{div}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\Bal}{\textrm{Bal}}
\newcommand{\Cut}{\textrm{Cut}}
\newcommand{\Ind}{\textrm{Ind}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Perp}{\perp \! \! \! \perp}

\newcommand{\Linv}{L^{\dagger}}
\newcommand{\tr}{\text{tr}}
\newcommand{\h}{\textbf{h}}
% \newcommand{\l}{\ell}
\newcommand{\x}{\textbf{x}}
\newcommand{\y}{\textbf{y}}
\newcommand{\bl}{\bm{\ell}}
\newcommand{\bnu}{\bm{\nu}}
\newcommand{\Lx}{\mathcal{L}_X}
\newcommand{\Ly}{\mathcal{L}_Y}
\DeclareMathOperator*{\argmin}{argmin}

\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\A}{\mathcal{A}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\N}{\mathbb{N}}
\newcommand{\E}{\mathcal{E}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden} 
\newtheorem{definition}{Definition}[section]

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}

\theoremstyle{remark}
\newtheorem{remark}{Remark}

\begin{document}

\title{Notes for Week of 12/24/18 - 12/31/18}
\author{Alden Green}
\date{\today}
\maketitle

\section{SETUP}
\paragraph{Data model.}

We are given two unknown distributions, $P$ and $Q$, both supported on $D \subset \Rd$, with continuous density functions $p$ and $q$, respectively. We have the capacity to sample from either distribution. Our goal is to test the nonparametric hypothesis $H_0: P = Q$ vs. the alternative $H_1: P \neq Q$.

Under the \textbf{binomial data model}, we sample data $\set{z_1, \ldots, z_n}$ as follows. For each $i$, we draw an independent Rademacher label $\ell_i \in \set{1, -1}$, $\mathbb{P}(\ell_i = 1) = \mathbb{P}(\ell_i = -1) = 1/2$. Then, if $\ell_i = 1$ we sample $z_i \sim P$, whereas if $\ell_i = -1$ we samle $z_i \sim Q$ otherwise. Define $1_X$ to be the length $n$ indicator vector for $\ell_i = 1$
\begin{equation*}
1_X(i) = 
\begin{cases}
1, \ell_i = 1\\
0 \text{ otherwise } 
\end{cases}
\end{equation*}
and similarly for $1_Y$
\begin{equation*}
1_Y(i) = 
\begin{cases}
1, \ell_i= -1 \\
0 \text{ otherwise } 
\end{cases}
\end{equation*}
Denote the number of positive labels $N = \sum_{i = 1}^{n} 1_x(i)$, and let $M$ be such that $n = N + M$. Then, let the \textbf{normalized label vector} $a$ be given by $a = \frac{1}{N}1_X - \frac{1}{M} 1_Y$.

\paragraph{Graph.}

Heuristically, our graph based test will attempt to identify whether samples from the same distribution are on average more similar than samples from different distributions. As a result, we introduce $K$, a \textbf{kernel function} which measures similarity, and make the following assumptions on $K$
\begin{itemize}
	\item $K: [0,\infty) \to [0,\infty)$ is non-increasing.
	\item The integral $\int_{0}^{\infty} K(r) r^d dr$ is finite.
\end{itemize}

and define

\begin{equation*}
K_{\epsilon}(z) = \frac{1}{\epsilon^d} K\left(\frac{z}{\epsilon}\right)
\end{equation*}

Given a sequence $\seq{\epsilon_n}$, form the \textbf{$\epsilon$-radius neighborhood graph} $G_{n,r} = (V_n,E_n)$ with $V_n = \set{z_1, \ldots, z_n}$ and $E_n = \set{(i,j): K_{\epsilon}(\norm{z_i - z_j}) > 0}$. Let $A$ be the adjacency matrix associated with $G_{n,r}$. Take $L_n = D - A$ to be the (unnormalized) \textbf{Laplacian matrix} of $A$ (where $D$ is the diagonal degree matrix with $D_{ii} = \sum_{j \in [n]} A_{ij}$). Denote by $B$ the $\abs{E} \times n$ \textbf{incidence matrix} of $A$, where we denote the $i$th row of $B$ as $B_i$ and set $B_i$ to have entry $A_{ij}$ in position $i$, $-A_{ij}$ in position $j$, and $0$ everywhere else.

\paragraph{Test Statistic.}
For a given neighborhood graph $G_{n,r}$, let $A$ be the adjacency matrix associated with $G_{n,r}$. Take $L_n = D - A$ to be the (unnormalized) \textbf{Laplacian matrix} of $A$ (where $D$ is the diagonal degree matrix with $D_{ii} = \sum_{j \in [n]} A_{ij}$). Denote by $B$ the $\abs{E} \times n$ \textbf{incidence matrix} of $A$, where we denote the $i$th row of $B$ as $B_i$ and set $B_i$ to have entry $A_{ij}$ in position $i$, $-A_{ij}$ in position $j$, and $0$ everywhere else. 
Now, we can define our \textbf{Laplacian smooth} test statistic
\begin{equation}
\label{eqn: laplacian_smooth_statistic}
T_2 := \left(\max_{\theta: \norm{B\theta}_2 \leq C_n} a^T \theta \right)^2
\end{equation}
for some sequence of positive numbers $\seq{C_n} \geq 0$.

\paragraph{Empirical Risk Minimization.}
To introduce the continuous limit of $T_2$, it will be useful to slightly recast the variational problem of (\ref{eqn: laplacian_smooth_statistic}) as an empirical risk minimization problem. (Really, all we are doing is introducing some new notation.) Let $\nu_n$ be the \textbf{empirical measure} induced by $\set{z_1, \ldots, z_n}$
\begin{equation*}
\nu_n \defeq \frac{1}{n} \sum_{i = 1}^{n} \delta_{z_i}
\end{equation*}

Then, for any mapping $u_n: \set{z_1, \ldots, z_n} \to \Reals$ such that $u_n \in L^2(\nu_n)$, let the \textbf{empirical risk functional} $R_n(u_n)$ be given by
\begin{equation}
\label{eqn: empirical_risk_functional}
R_n(u_n) \defeq - \sum_{i = 1}^{N} u_n(z_i) \widetilde{\ell}_n(z_i)
\end{equation}
where $\widetilde{\ell}_n: \set{z_1, \ldots, z_n} \to \set{0,1}$ is the \textbf{normalized label function} defined by $\widetilde{\ell}_n(z_i) := a_i$.

To relate the risk functional of (\ref{eqn: empirical_risk_functional}) to the variational problem of (\ref{eqn: laplacian_smooth_statistic}), we introduce the \textbf{constrained empirical risk functional}, $R_n^{(con)}(u_n)$, defined by
\begin{equation*}
R_n^{(con)}(u_n) := 
\begin{cases}
R_n(u_n), \text{ if $\mathcal{E}_n^2(u_n) \leq 1$} \\
\infty, \text{ otherwise }
\end{cases}
\end{equation*}
where $\mathcal{E}_n^2(u_n)$ is the \textbf{Laplacian regularization functional} given by
\begin{equation*}
\mathcal{E}_n^2(u_n) \defeq \frac{1}{n^2 \epsilon_n^{d+1}} \sum_{i = 1}^{n} \sum_{j = 1}^{n} K\left(\frac{\norm{z_i - z_j}}{\epsilon_n}\right)(u_n(z_i) - u_n(z_j))^2.
\end{equation*}

Note that for $B$ the incidence matrix of the $\epsilon$-radius neighborhood graph $G_{n,\epsilon_n}$ as defined above, and for $\theta = \set{u_1(z_i), \ldots, u_n(z_i)}$, we have
\begin{equation*}
N^2 \epsilon_N^{d+1} \mathcal{E}_N^2(u_N) = \norm{B \theta}_2, ~~~ \mathcal{E}_N^2(u_N) \leq 1 \Leftrightarrow \norm{B \theta}_2 \leq N^2 \epsilon_N^{d+1}
\end{equation*}
and $a^T \theta = - R_n(u_n)$.
As a result, letting $C_n = N^2 \epsilon_N^{d+1}$, we have that for
\begin{equation}
\label{eqn: minimizer_of_constrained_empirical_risk_functional}
u_n^{\star} := \argmin_{u_n \in L^2(\nu_n)} R_n^{(con)}(u_n)
\end{equation}
the following relation holds:
\begin{equation*}
T_2^{1/2} = R_n(u_n^{\star})
\end{equation*}

\paragraph{Continuum limit.}

As the preceding manipulations make clear, the statistic $T_2$ can be seen as a constrained minimization problem with constraint enforced by a regularization functional over the neighborhood graph $G_{n,\epsilon}$. It is well known that, for an appropriate schedule of $\seq{\epsilon_n}$ and data generated from a density satisfying certain regularity conditions, such regularization functionals are well behaved in the limit.

Let $\nu = \frac{p + q}{2}$. For $u \in L^2(\nu)$, define the \textbf{continuous risk functional} $R(u)$ via
\begin{equation*}
R(u) = -\int_{D} u(x) \bigl(p(x) - q(x) \bigr) dx
\end{equation*}
the \textbf{weighted $L^2$ regularization functional}
\begin{equation*}
\mathcal{E}_{\infty}^2(u) = \int_D \norm{\nabla u(x)}^2 \mu^2(x) dx
\end{equation*}
and the \textbf{constrained continuous risk functional} $R^{(con)}(u)$ as
\begin{equation*}
R^{(con)}(u) =
\begin{cases}
R(u), \text{if $\mathcal{E}_{\infty}^2(u) \leq 1$ } \\
\infty, \text{otherwise}
\end{cases}
\end{equation*}

Let $u^{\star}$ be defined analogously to $u_n^{\star}$,
\begin{equation}
\label{eqn: minimizer_of_constrained_continuous_risk_functional}
u^{\star} = \argmin_{u \in L^2(\nu)} R^{(con)}(u)
\end{equation}

\section{RESULTS}
\begin{theorem}
	\label{thm: consistency_of_empirical_risk_minimizer}
	Consider a sequence $\seq{\epsilon_n} \to 0$ satisfying
	\begin{equation*}
	\left(\frac{\log n}{n}\right)^{1/d} = o(\epsilon_n).
	\end{equation*}
	Then, for $u_n^{\star}$ satisfying (\ref{eqn: minimizer_of_constrained_empirical_risk_functional}) and likewise $u^{\star}$ satisfying (\ref{eqn: minimizer_of_constrained_continuous_risk_functional}), with probability one:
	\begin{equation}
	R_n(u_n^{\star}) \to R(u^{\star})
	\end{equation}
\end{theorem}

\begin{proof}[Proof attempt of Theorem \ref{thm: consistency_of_empirical_risk_minimizer}]

We begin by restricting ourselves to the case where $\underset{n \to \infty}{\limsup}~ \E_n^2(u_n^{\star}) \leq 1$.
As a result of Lemma 
\end{proof}
\subsection{Technical results.}

Lemma \ref{lem: conv_of_risk_functional} is very similar to Proposition 2.7 in \textcolor{red}{cite (Garcia-Trillos 16)}.  
\begin{lemma}
	\label{lem: conv_of_risk_functional}
	With probability one the following statement holds: Let $\seq{u_n}$ be a sequence of $[-1,1]$-valued functions, with $u_n \in L^1(\nu_n)$. If $u_n \overset{TL^1}{\to} u$ as $n \to \infty$, then
	\begin{equation*}
	\lim_{n \to \infty} R_n(u_n) = R(u).
	\end{equation*}
\end{lemma}

Lemma \ref{lem: weak_limit_of_label_function} is needed to prove Lemma \ref{lem: conv_of_risk_functional}. It is very similar to a result from \textcolor{red}{cite (Garcia-Trillos 16)}. To understand Lemma \ref{lem: weak_limit_of_label_function}, we must introduce a notion of weak convergence of functions.

\begin{definition}
	Given a sequence of functions $g_n \in L^1(\nu)$, and $g \in L^1(\nu)$, we say \textbf{$g_n$ converges weakly to $g$}, $g_n \rightharpoonup g$ if for all $L^{\infty}(\nu)$,
	\begin{equation*}
	\lim_{n \to \infty} \int_D g_n(x) f(x) d\nu(x) = \int_D g(x) f(x) d\nu(x)
	\end{equation*}
	Given a sequence $\seq{u_n}$ with $u_n \in L^1(\nu_n)$, we say that $\seq{u_n}$ \textbf{converges weakly} to $u \in L^1(\nu)$, $u_n \rightharpoonup u$,  if the sequence of functions $\set{u_n \circ T_n} \in L^1(\nu)$ converges weakly to $u$, for $T_n$ stagnating transportation maps.
\end{definition}

\begin{lemma}
	\label{lem: weak_limit_of_label_function}
	For the label function $\ell_n: \set{z_1, \ldots, z_n} \to \set{-1,1}$ defined by
	\begin{equation}
	\label{eqn: label_function}
	\ell_n(z_i) := \ell_i, ~~~ i \in 1,\ldots,n
	\end{equation}
	with probability one $\ell_n \rightharpoonup \frac{p - q}{2 \mu}$. 
\end{lemma}

\begin{lemma}
	\label{lem: gamma_convergence_of_energy_function}
	The Laplacian regularization functional $\E_n^2(u_n)$ satisfies the following three properties, 
\end{lemma}

\section{PROOFS}
\begin{proof}[Proof of Lemma \ref{lem: conv_of_risk_functional}]
	We begin by removing the effect of the random normalization in $\widetilde{\ell}_n(z_i)$ via
	\begin{equation*}
	- R_n(u_n) = \frac{2}{n} \sum_{i = 1}^{n} u_n(z_i) \ell_n(z_i) + \sum_{i = 1}^{n} \left(\frac{2}{n} \ell_n(z_i) - \widetilde{\ell}_n(z_i) \right)u_n(z_i)
	\end{equation*}
	First, we show the second term converges to zero with probability one,
	\begin{align*}
	\abs{\sum_{i = 1}^{n} \left(\frac{2}{n} \ell_n(z_i) - \widetilde{\ell}_n(z_i) \right)u_n(z_i)} & \leq \sum_{i = 1}^{n} \abs{\frac{2}{n} \ell_n(z_i) - \widetilde{\ell}_n(z_i)} \abs{u_n(z_i)} \\
	& \leq \left(\abs{\frac{1}{N} - \frac{1}{n/2}} + \abs{\frac{1}{M} - \frac{1}{n/2}}\right) \sum_{i = 1}^{n} \abs{u_n(z_i)} \\
	& \leq \left(\abs{\frac{N - n/2}{N}} + \abs{\frac{M - n/2}{M}}\right) \cdot \frac{2}{n} \sum_{i = 1}^{n}\abs{u_n(z_i)}
	\end{align*}
	Then, the $TL^1$ convergence of $u_n$ to $u \in L^1(\nu)$ implies
	\begin{equation*}
	\frac{2}{n} \sum_{i = 1}^{n}\abs{u_n(z_i)} \overset{n}{\to} 2 \norm{u}_{L^1(\nu)}
	\end{equation*}
	and by standard concentration results of binomial random variables, with probability one
	\begin{equation*}
	\abs{\frac{N - n/2}{N}}, \abs{\frac{M - n/2}{M}} \overset{n}{\to} 0,
	\end{equation*}
	and so with probability one
	\begin{equation*}
	\abs{\sum_{i = 1}^{n} \left(\frac{2}{n} \ell_n(z_i) - \widetilde{\ell}_n(z_i) \right)u_n(z_i)} \overset{n}{\to} 0.
	\end{equation*}

	Now, we rewrite the first term in the summand on the right hand side using transportation maps,
	\begin{align*}
	\frac{2}{n} \sum_{i = 1}^{n} u_n(z_i) \ell_n(z_i) & = 2 \int_D u_n(z) \ell_n(z) d\nu_n(z) \\
	& \overset{(i)}{=} 2 \int_D  \bigl(u_n \circ T_n (z) \bigr) \bigl( \ell_n \circ T_n (z)  \bigr) d\nu(z) \\
	& = 2 \int_D  \bigl(u_n \circ T_n (z)  - u(z) \bigr) \bigl( \ell_n \circ T_n (z)  \bigr) d\nu(z) + 2 \int_D  \bigl(u(z) \bigr) \bigl( \ell_n \circ T_n (z)  \bigr) d\nu(z) \\
	\end{align*}
	with $(i)$ following from the change of variables formula
	\begin{equation*}
	\int_D f(T(x)) d \theta(x) = \int_D f(z) d T_{\sharp} \theta x
	\end{equation*}
	where $f: D \to \Reals$ is an arbitrary Borel function, $\theta$ a Borel measure, and $T_{\sharp} \theta$ the push-forward measure of $\theta$.

	Now, the first term converges to zero,
	\begin{equation*}
	\abs{2 \int_D  \bigl(u_n \circ T_n (z)  - u(z) \bigr) \bigl( \ell_n \circ T_n (z)  \bigr) d\nu(z)} \leq 2 \int_D \abs{u_n \circ T_n (z)  - u(z)} d\nu(z) \overset{n}{\to} 0.
	\end{equation*}
	by the boundedness of $\ell_n$ and the $TL^1$ convergence of $u_n$ to $u$.

	By Lemma \ref{lem: weak_limit_of_label_function}, with probability one the 2nd term converges to (negative of) the risk functional,
	\begin{equation*}
	2 \int_D  \bigl(u(z) \bigr) \bigl( \ell_n \circ T_n (z)  \bigr) d\nu(z) \overset{n}{\to} \int_D u(z) \left(\frac{p(z) - q(z)}{\mu(z)}\right) d\nu(z) = -R(u).
	\end{equation*}
\end{proof}

\begin{proof}[Proof of Lemma \ref{lem: conv_of_risk_functional}]
	Lemma \ref{lem: weak_limit_of_label_function} is essentially a restatement of Lemma 2.5 from \textcolor{red}{cite (Garcia-Trillos 16)}. For completeness purposes, we restate that result here.

	\begin{lemma}
		\label{lem: weak_conv_label_function}
		Let $f(z) = \mathbb{E}(\ell \vert Z = z)$ be the conditional expectation of $\ell$ given $Z = z$.  
		For the label function $\ell_n: \set{x_1, \ldots, x_n} \to \set{-1,1}$ defined as in (\ref{eqn: label_function}), with probability one $\ell_n \rightharpoonup f$.
	\end{lemma}
	Given Lemma \ref{lem: weak_conv_label_function}, all that is needed to show Lemma \ref{lem: weak_limit_of_label_function} is that $f = \frac{p - q}{2\mu}$. This follows from a simple application of Bayes Rule. 
\end{proof}

\end{document}

### 2019-01-30 notes/2018.11.7.LaplacianSmoothingTest.tex
added backlog of notes

--- 2018-12-19 notes/graphtestingmasternotes.tex
@@
 \newcommand{\norm}[1]{\left\lVert#1\right\rVert}
-\newcommand{\dotp}[2]{\langle #1 , #2 \rangle}
 \newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
@@
 \newcommand{\set}[1]{\left\{#1\right\}}
-\newcommand{\Perp}{\perp \! \! \! \perp}
 
@@
 \newcommand{\tr}{\text{tr}}
-\newcommand{\h}{\textbf{h}}
-% \newcommand{\l}{\ell}
-\newcommand{\x}{\textbf{x}}
-\newcommand{\y}{\textbf{y}}
-\newcommand{\Lx}{\mathcal{L}_X}
-\newcommand{\Ly}{\mathcal{L}_Y}
 \DeclareMathOperator*{\argmin}{argmin}
@@
 \newcommand{\F}{\mathcal{F}}
-\newcommand{\X}{\mathcal{X}}
-\newcommand{\Rd}{\Reals^d}
 
@@
 
-\title{Graph Testing Notes}
+\title{Laplacian smooth test statistic for two-sample testing}
 \author{Alden Green}
@@
 
+\section{Goals}
+\begin{itemize}
+	\item Find an asymptotic null distribution.
+\end{itemize}
+
 \section{Setup}
 
-\paragraph{Data model.}
+We observe data $X_1, \ldots, X_n \sim P$ and $Y_1, \ldots, Y_m \sim Q$. Our goal is to test the hypothesis $H_0: P = Q$ vs. the alternative $H_a: P \neq Q$.
 
-We are given two distributions, $P$ and $Q$, with the ability to sample from either one. Our goal is to test the hypothesis $H_0: P = Q$ vs. the alternative $H_a: P \neq Q$.
-
-Under the \textbf{binomial data model}, our sampling procedure is to draw i.i.d Rademacher labels $L_i \in \set{1, -1}$ for $i \in \set{1, \ldots, N}$, and then sample 
-$Z_i \sim P$ if $L_i = 1$ and $Z_i \sim Q$ otherwise. Define $1_X$ to be the length-$N$ indicator vector for $L_i = 1$
+Let $Z  = (X_1, \ldots, X_n, Y_1, \ldots, Y_m)$. Define $1_X$ to be the $n + m$ length indicator vector for $X$
 \begin{equation*}
@@
 \begin{cases}
-1, L_i = 1\\
+1, i \in 0, 1, \ldots, n \\
 0 \text{ otherwise } 
@@
 \begin{cases}
-1, L_i= -1 \\
+1, j \in n + 1, \ldots, m \\
 0 \text{ otherwise } 
@@
 \end{equation*}
-and define $a = \frac{1_X}{N / 2} - \frac{1_Y}{N / 2}$.
+and define $a = \frac{1_X}{m} - \frac{1_Y}{n}$.
 
-Under the \textbf{fixed label data model} we use the same data generating process as above, except fix $\mathcal{L}_X = \set{1, \ldots, N/2}$ and $\mathcal{L}_Y = \set{N/2, \ldots, N}$.
-Say that $L_i = 1$ for $i \in \mathcal{L}_X$ and $L_i = -1$ for $i \in \mathcal{L}_Y$, and call $\set{ X_1, \ldots, X_{\abs{\Lx}} } = \set{Z_i: i \in \mathcal{L}_X}$ and likewise for $Y$.
+Form an $n + m \times n + m$ similarity matrix $A$, where $A_{ij} = K(Z_i, Z_j)$ for some unspecified choice of $K$, and take $L = D - A$ to be Laplacian matrix of $A$ (where $D$ is a diagonal matrix with $D_{ii} = \sum_{j \in [n + m]} A_{ij}$).
 
-\paragraph{Graph.}
-
-Form an $N \times N$ Gram matrix $A$, where $A_{ij} = K(Z_i, Z_j)$ for \textbf{kernel function} $K$.
-Let $G = (V,E)$ with $V = \set{Z_1, \ldots, Z_n}$ and $E = \set{A_{ij}: 1 \leq i < j \leq n}$.
-Take $L = D - A$ to be the (unnormalized) \textbf{Laplacian matrix} of $A$ (where $D$ is the diagonal degree matrix with $D_{ii} = \sum_{j \in [n + m]} A_{ij}$).
-Denote by $B$ the $M \times N$ \textbf{incidence matrix} of $A$, where we denote the $i$th row of $B$ as $B_i$ and set $B_i$ to have entry $A_{ij}$ in position $i$, $-A_{ij}$ in position $j$, and $0$ everywhere else.
-
-\paragraph{Resistance distances.}
-
-There are many distances one can define over nodes in a graph. The \textbf{resistance distance between nodes $u$ and $v$}, $R_{uv}$, is defined as
+We are ready to define our test statistic. 
 \begin{equation*}
-R_{uv} = (e_u - e_v)^T \Linv (e_u - e_v).
+T_2^2 = \left(\max_{\theta: \theta^T L \theta \leq 1} a^T \theta. \right)^2
 \end{equation*}
 
-\paragraph{Test statistics.}
+\paragraph{Spectral properties of $L$.}
 
-We begin by defining our \textbf{laplacian smooth} test statistic. 
-\begin{equation*}
-T_2 = \left(\max_{\theta: \norm{B\theta}_2 \leq 1} a^T \theta \right)^2 = a^T \Linv a. 
-\end{equation*}
+Define the pseudo-inverse of $L$ to be $\Linv$.
+In what follows, we will assume $A$ defines a connected graph $G$.
+In this setting, it is well known that $L$ has exactly one $0$ eigenvalue, with corresponding eigenvector ${\bf 1}$.
+Let $P_{1^{\perp}}$ be the projection onto the linear subspace orthogonal to this eigenvector.
 
-(Bhattacharya 2018) defines a general notion of 2-sample \textbf{graph-based test statistics}
-\begin{equation*}
-T_{\mathcal{G}} = \frac{1}{N^2}\sum_{i = 1}^{n} \sum_{j = n + 1}^{n + m} A_{ij}
-\end{equation*}
-Although he develops theory for this statistic in the context of $k$NN and minimum spanning tree graphs, we will at present consider it for the complete weighted similarity graph defined by $A$ above.
-Then, we can write
-\begin{equation*}
-T_{\mathcal{G}} = a^T L a.
-\end{equation*}
+\paragraph{Poissonization.}
 
-Finally, define $\mathcal{H}$ to be a \textbf{reproducing kernel Hilbert space} with $K$ the associated kernel.
-Let $\mathcal{F}$ be the unit ball of $\mathcal{H}$, and let the evaluation of $f \in \mathcal{F}$ at the sample points $Z_1, \ldots, Z_N$ be denoted by $\textbf{f} = f(Z_1, \ldots, Z_N)$.
-Then, the statistic $\text{MMD}_b$ of (Gretton 2012) can be written as
-\begin{equation*}
-T_{\mathcal{K}} = \sup_{f \in \mathcal{F}} a^T \textbf{f} = a^T K a. 
-\end{equation*}
+For $p \in (0,1)$, draw $U_1, \ldots, U_{N} \sim Bern(p)$.
+Then, draw $Z_i \sim P_{2 U_i - 1}$.
+Consider $a = (a_i)_{i = 1}^{N}$ with $a_i = 2 U_i - 1$.
+Let the null hypothesis be $H_0: P_1 = P_2$ and the alternative be $H_a: P_1 \neq P_2$.
 
@@
 
-We will need distances between probability measures for two different purposes.
-The first is that they are self-evidently useful in analyzing limiting distributions of statistics (in particular in this case, our test statistics).
-
 For a function $f$, define its \textbf{Lipschitz norm} $\norm{f}_{L}$ to be
 \begin{equation*}
-\inf \set{K : \abs{f(x) - f(y)} \leq K \norm{x - y}}.
+\inf K : \abs{f(x) - f(y)} \leq K \norm{x - y}.
 \end{equation*}
@@
 
-The second reason we will use distances between probability measures is that they themselves make for good test statistics!
+The following lemma allows us to translate from an upper bound on Wasserstein distance to Kolmogorov distance.
+\begin{lemma}[Wasserstein to Kolmogorov distance]
+	\label{lem: wass_to_ks_distance}
+	For any probability measures $\mu$, $\nu$ with corresponding cdfs $F_{\mu}$ and $F_{\nu}$ and any $\epsilon' > 0$, there exists some $\epsilon > 0$ such that
+	\begin{equation*}
+	\mathcal{W}(\mu, \nu ) < \epsilon \implies \sup_{t} \abs{F_{\mu}(t) - F_{\nu}(t)} \leq \epsilon'.
+	\end{equation*}
+\end{lemma}
 
-An \textbf{integral probability metric} (IPM) with respect to a function class $\F$ is defined
+\section{Related Work}
+
+(Bhattacharya 2018) defines a general notion of 2-sample graph-based test statistics
+\begin{definition}
 	\begin{equation*}
-\underset{f \in \F}{\sup} ~ \Expect{f(X)} - \Expect{f(Y)}
+	T(G) = \sum_{i = 1}^{n} \sum_{j = n + 1}^{n + m} 1( e_{ij} \in E)
 	\end{equation*}
-for $X \sim P$, $Y \sim Q$.
-
-Hereafter, we will assume $P$ and $Q$ are absolutely continuous with respect to Lebesgue measure, with density functions $p$ and $q$, respectively.
-Denote the \textbf{mixture density} by $\mu = \frac{p + q}{2}$.
-
-Denote the \textbf{gradient} of a function $f$ by $\nabla_x$.
-Then we can define the \textbf{Sobolev semi-norm} and \textbf{dot product}, $\norm{f}_{W_0^{1,2}(\X, \mu^2)}$ and $\dotp{f}{g}_{W_0^{1,2}(\X, \mu^2)}$, by
+\end{definition}
+and letting $a_0 = \frac{1}{2} (1_X - 1_Y)$, we can write
 \begin{equation*}
-\dotp{f}{g}_{W_0^{1,2}(\X, \mu)} = \int_{\X} \dotp{\nabla_x f(x)}{\nabla_x g(x)}_{\Rd} \mu^2(x), ~~ \norm{f}_{W_0^{1,2}(\X, \mu)} = \sqrt{\int_{\X} \norm{\nabla_x f(x)}^2 \mu^2(x) dx }
+T(G) = a_0^T L a_0.
 \end{equation*}
 
-Let the \textbf{Sobolev space}, $W^{1,2}(\mathcal{X}, \mu^2)$, be 
+(Gretton 2012) considers the test statistic
 \begin{equation*}
-W^{1,2}(\mathcal{X},\mu^2) = \set{f: \X \to \Reals, \int_{\X} \norm{\nabla_x f(x)}^2 \mu^2(x) dx < \infty}.
+T = a^T A a.
 \end{equation*}
-and denote by $W_0^{1,2}(\mathcal{X}, \mu^2)$ the restriction of $W^{1,2}(\mathcal{X}, \mu^2)$ to functions which vanish at the boundary of $\X$.
-Note that $\norm{f}_{W_0^{1,2}(\X, \mu^2)}$ defines a semi-norm over $W_0^{1,2}(\X, \mu^2)$.
-Finally, let $B_W(\X, \mu^2)$ be the \textbf{unit ball} of $W_0^{1,2}(\X, \mu^2)$, meaning
-\begin{equation*}
-B_W(\X, \mu^2) = \set{f \in W_0^{1,2}(\X, \mu^2): \norm{f}_{W_0^{1,2}(\X, \mu^2)} \leq 1}
-\end{equation*}
-
-Now we can define the \textbf{Sobolev IPM}, $\mathcal{S}_{\mu^2}(P,Q)$ It is simply an IPM where the function class is the Sobolev unit ball with respect to $\mu^2$.
-\begin{equation*}
-\mathcal{S}_{\mu^2}(P,Q) \overset{\mathrm{def}}{=} \sup_{f \in B_W} \biggl\{ \Expect{f(X)} - \Expect{f(Y)}\biggr\}
-\end{equation*}
-
-\paragraph{Holder functions.}
-We will show that the Laplacian constraint $\norm{B \theta}_2 \leq 1$ is very similar to the constraint $f_{\theta} \in B_W(X,\mu^2)$ for the right choice of $K$, over all Holder functions.
-
-For mapping $f: \Rd \to \Reals$ and $\beta$ a positive integer, we say $f$ is a \textbf{$\beta$-Holder function} if there exists $C > 0$ such that for all $x, y \in \X$
-\begin{equation*}
-\abs{f^{(\beta - 1)}(x) - f^{(\beta - 1)}(y)} \leq K \norm{x - y}
-\end{equation*}
-Roughly speaking, this means the functions have bounded $\beta$ partial derivatives.
 
@@
 
-Conjectures \ref{conj: spectral_measure_conv} and \ref{conj: diagonal_entries} will be needed for Theorem \ref{thm: asymptotic_null_distribution}.
+The following will be needed for Theorem \ref{thm: asymptotic_null_distribution}.
 
@@
 
-\section{DESIRED RESULTS}
-
-\begin{theorem}
-	\label{thm:convprob_of_T2}
-	For bandwidth parameter $h >0$ and decreasing function $k(\cdot,\cdot)$, write
-	\begin{equation*}
-	K(Z_i,Z_j) = \frac{1}{h^m} k(\norm{Z_i - Z_j}^2 / h^2).
-	\end{equation*}
-
-	For Sobolev IPM $\mathcal{S}_{\mu^2}(P,Q)$ as defined above,
-	\begin{equation*}
-	\sqrt{T_2} \convprob \mathcal{S}_{\mu^2}(P,Q)
-	\end{equation*}
-\end{theorem}
-
-\begin{proof}[Proof attempt of Proposition \ref{thm:convprob_of_T2}]
-	Recall that, for incidence matrix $B$, 
-	\begin{equation*}
-	\sqrt{T_2} = \left(\max_{\theta: \norm{B\theta}_2 \leq 1} a^T \theta \right).
-	\end{equation*}
-
-	We expand $\abs{\sqrt{T_2} - \mathcal{S}_{\mu^2}(P,Q)}$,
-	\begin{align}
-	\label{eqn: bias_variance_expansion}
-	\abs{\sqrt{T_2} - \mathcal{S}_{\mu^2}(P,Q)} & \leq \abs{\max_{\theta: \norm{B\theta}_2 \leq 1} \bigl\{a^T \theta \bigr\} - \sup_{f \in B_W(\X, \mu^2)} \left\{ \mathbb{P}_n(f) - \mathbb{Q}_n(f) \right\} } \nonumber \\
-	& + \abs{\sup_{f \in B_W(\X, \mu^2)} \left\{ \mathbb{P}_n(f) - \mathbb{Q}_n(f) \right\} - \sup_{f \in B_W(\X, \mu^2)} \left\{ \mathbb{P}(f) - \mathbb{Q}(f) \right\} }
-	\end{align}
-
-	\textcolor{red}{(The following statement would hold only if Proposition \ref{prop: sobolev_donsker} held over $B_W(\X, \mu^2)$, rather than over $B_W([0,1], \lambda)$ for $\lambda$ Lebesgue measure.)}
-
-	By Proposition \ref{prop: sobolev_donsker}, the second term in the summand on the right hand side of (\ref{eqn: bias_variance_expansion}) is $o_P(1)$.
-
-	\textcolor{red}{(The following statement would hold only if Proposition \ref{prop: convprob_regularization_functional} were uniform over $B_W(\X, \mu^2)$ rather than over the class of $\alpha$-Holder functions $\F_{\alpha}$)}
-
-	Then, Proposition \ref{prop: convprob_regularization_functional} implies that for any $\epsilon > 0$, there exists $N$ such that for $n \geq N$, 
-	\begin{equation*}
-	\sup_{f \in B_W(\X, \mu^2)} \left\{ \mathbb{P}_n(f) - \mathbb{Q}_n(f) \right\} - \max_{\theta: \norm{B\theta}_2 \leq 1} \bigl\{a^T \theta \bigr\} \leq \epsilon
-	\end{equation*}
-	with high probability.
-
-	To complete the proof, we will have to show that for any $\epsilon > 0$, there exists $N$ such that for $n \geq N$, 
-	\begin{equation*}
-	\max_{\theta: \norm{B\theta}_2 \leq 1} \bigl\{a^T \theta \bigr\} - \sup_{f \in B_W(\X, \mu^2)} \left\{ \mathbb{P}_n(f) - \mathbb{Q}_n(f) \right\} \leq \epsilon
-	\end{equation*}
-	with high probability.
-\end{proof}
-
 \section{Results}
-
-\paragraph{Expectation of two-sample test statistics.}
-
-The expectation of the above statistics is potentially a good way to understand their large sample behavior, as quadratic forms often satisfy laws of large numbers assuming the matrices are well-conditioned.
-
-\begin{proposition}
-	\label{prop: expectation_of_TG}
-	Draw $Z$ and $a$ under the binomial data model, and assume both $P$ and $Q$ are absolutely continuous with respect to Lebesgue measure over Euclidean space $\Reals^d$.
-Write $h_0(x) = p(x) - q(x)$ with empirical analogue $\mathbf{h_0} = (h_0(Z_1), \ldots, h_0(Z_n) )$.
-Then
-	\begin{align}
-	\label{eqn: expectation_of_TG}
-	\Expect{T_{\mathcal{G}}} = & \int \int K(\norm{\x - \y}) \left[p(\x) + q(\x)\right] \left[p(\y) + q(\y)\right] d\x d\y \nonumber \\
-	- & \frac{N}{N - 1} \int \int K(\norm{\x - \y}) \left[h_0(\x)\right]^2 \frac{p(\y) + q(\y)}{p(\x) + q(\x)} d\x d\y \nonumber \\
-	+ & \frac{N}{N - 1} \int \int K(\norm{\x - \y}) \left[h_0(\x) - h_0(\y)\right]^2 \left[p(\y) + q(\y)\right] \left[p(\x) + q(\x)\right] d\x d\y.
-	\end{align}
-\end{proposition}
-
-Note that even under the null hypothesis, where $h_0 = 0$,
-\begin{equation*}
-\Expect{T_{\mathcal{G}}} = \int \int K(\norm{\x - \y}) \left[p(\x) + q(\x)\right] \left[p(\y) + q(\y)\right] d\x d\y
-\end{equation*}
-which is not distribution-free, unlike in the case of the $k$NN graph.
-
-Another interesting consequence of Proposition \ref{prop: expectation_of_TG} comes when we take $K(\x,\y) = K(\frac{\norm{\x - \y}}{t})$ and let $t \to 0$.
-
-\begin{proposition}
-	\label{prop: expectation_of_TG_with_kde}
-	If $p$ and $q$ are Lipschitz continuous functions with bounded hessians, and $K$ is a continuous function on $R^+$ such that $x^{2 + d} K(x) \in L_2$, then under the same setup as in Proposition \ref{prop: expectation_of_TG},
-	\begin{equation}
-	\label{eqn: expectation_of_TG_with_kde}
-	\frac{N - 1}{N}\underset{t \to 0}{\lim} \frac{1}{t^d} \Expect{T_{\mathcal{G}}} = \int  h_0(\x)^2 d\x + \int (p(\x) + q(\x))^2 d\x
-	\end{equation}
-\end{proposition}
-
-We turn now to the expectation of the Laplacian smooth statistic.
-\begin{proposition}
-	\label{prop: expectation_of_T2}
-	Under the fixed label data model
-	\begin{equation}
-	\label{eqn: expectation_of_T2}
-	\Expect{a^T \Linv a} = \Expect{R_{X_1 Y_1}} - \frac{N - 1}{2N} \Expect{R_{X_1,X_2}} - \frac{N - 1}{2N} \Expect{R_{Y_1, Y_2}}
-	\end{equation}
-\end{proposition}
-
-Note that, although the resistance distances are between only two nodes, in each case the expectation is over the entire (random) graph $G$.
-
-\paragraph{Asymptotic null distribution for $T_2$.}
-
-We can compute an asymptotic null distribution for $T_2$, although its formulation depends on the eigenvalues of the matrix $\Linv$ which themselves are not obvious.
-
-\begin{theorem}
-	\label{thm: asymptotic_null_distribution}
-	Denote the scaled version of the Laplacian smooth test statistic
-	\begin{equation*}
-	W_n = \sqrt{\frac{N^4}{32 \cdot \tr((\Linv)^2)}}\Bigl(T_2^2 - \frac{\tr(\Linv)}{4N^2} \Bigr).
-	\end{equation*}
-
-	If Conjectures \ref{conj: spectral_measure_conv} and \ref{conj: diagonal_entries} hold,
-	\begin{equation*}
-	\lim_{n \to \infty} \sup_{t} \abs{\Prob{W_n \leq t} - \Phi(t)} = 0.
-	\end{equation*}
-\end{theorem}
-
-To prove Theorem \ref{thm: asymptotic_null_distribution}, we will need the following calculations of moments under $H_0$.
-
-\begin{lemma}
-	\label{lem: cond_expectation_of_T2}
-	Under $H_0$, the conditional expectation $\Expect{T_2 \vert Z} = \frac{\tr(\Linv)}{N^2}$. 
-\end{lemma}
-
-\begin{lemma}
-	\label{lem: cond_var_of_T2}
-	Under $H_0$, the conditional variance $\Var{T_2 \vert Z} = \frac{32 \tr[(\Linv)^2]}{N^4}$. 
-\end{lemma}
-
-\section{Supplemental Results}
-
-\paragraph{Empirical process over Sobolev classes.}
-
-The following theorem is a stand-in; it handles only functions with domain on the unit interval, and is stated specifically with respect to Lebesgue measure.
-\begin{proposition}
-	\label{prop: sobolev_donsker}
-	Let $\F$ be the set of all absolutely continuous functions $f: [0,1] \to \Reals$ such that $\norm{f}_{\infty} \leq 1$ such that $\int (f'(x))^2 dx \leq 1$.
-Then, there exists a constant $K$ such that for every $\epsilon > 0$,
-	\begin{equation*}
-	\log N_{[]}(\epsilon, \mathcal{F}, \norm{\cdot}_{\infty}) \leq K\left(\frac{1}{\epsilon}\right).
-	\end{equation*}
-
-	Thus, the class $\mathcal{F}$ is $P$-Donsker (and $P$-Glivenko-Cantelli) for all $P$.
-\end{proposition}
-
-\paragraph{Regularization functionals.}
-
-When taking the supremum over functions which satisfy $\norm{B \theta}_2 \leq 1$, we will argue that this constraint is well-behaved in the limit, i.e.
-that it converges to the \textbf{regularization functional} $\norm{\cdot}_{W_0^{1,2}(\X, \mu^2)}$.
-Proposition \ref{prop: convprob_regularization_functional} makes this convergence uniform over the set of $3$-Holder functions (essentially functions with bounded $3rd$ derivative).
-Proposition \ref{prop: bousquet04} makes this convergence only pointwise, but merely requires that $f$ have bounded 2nd derivative.
-
-\begin{proposition}
-	\label{prop: convprob_regularization_functional}
-	Let $\mathcal{F}_{\alpha}$ be a unit ball in the space of $\alpha$-Holder functions, and define $k(\cdot,\cdot)$ as in Theorem \ref{thm:convprob_of_T2}.
-For function $f \in \F_{\alpha}$, denote $f$ evaluated on the data, $\mathbf{f} = (f(Z_1), \ldots, f(Z_N))$.
-Then, there exists a constant $c$ depending only on $k$ such that for $\alpha \geq 3$ and a sequence $(h_n) \to 0$ such that
-	\begin{equation*}
-	\sup_{f \in \mathcal{F}_{\alpha}} \abs{ \norm{B \mathbf{f}_2} - \norm{f}_{W_0^{1,2}(\X, \mu^2)} } \convprob 0
-	\end{equation*}
-\end{proposition}
-
-\begin{proposition}[Bousquet 04]
-	\label{prop: bousquet04}
-	If $p$ and $q$ are Lipschitz continuous functions with bounded hessians, and $K$ is a continuous function on $R^+$ such that $x^{2 + d} K(x) \in L_2$, then 
-	\begin{align}
-	\label{eqn: bousquet04}
-	& \underset{t \to 0}{\lim} \frac{d}{Ct^{d+2}} \int  K(\norm{\x - \y}/ t)(h_0(\x) - h_0(\y))^2 (p(\x) + q(\x))(p(\y) + q(\y)) d\x d \y \nonumber \\
-	& = \int \norm{\nabla h_0(\x)}^2 (p(\x) + q(\x))^2 d\x
-	\end{align}
-\end{proposition}
-
-\begin{lemma}[von Luxburg 12]
-	\label{lem: vonluxburg12}
-	Assume $P$ and $Q$ are absolutely continuous with respect to Lebesgue measure on Euclidean space $\Reals^d$, with density functions $p$ and $q$, respectively.
-Let $K(x,y) = \frac{1}{(2\pi \sigma^2)}^{d/2} \exp{-\frac{\norm{x - y}^2}{2 \sigma^2}}$.
-
-	Under some regularity assumptions on $p$ and $q$, if $n \to \infty$, $\sigma \to 0$ , and $n\sigma^{d+2}/ \log(n) \to \infty$, then
-	\begin{align*}
-	n R_{XY} \to \frac{2}{p(X) + q(X)} + \frac{2}{p(Y) + q(Y)} \textit{ almost surely }
-	\end{align*}
-	with equivalent statements holding for $X_1, X_2$ and $Y_1, Y_2$.  
-\end{lemma}
 
@@
 
-\paragraph{Translating from Wasserstein to Kolmogorov distance.}
+\paragraph{Analytic form for $T_2^2$.}
 
-\begin{lemma}[Wasserstein to Kolmogorov distance]
-	\label{lem: wass_to_ks_distance}
-	For any probability measures $\mu$, $\nu$ with corresponding cdfs $F_{\mu}$ and $F_{\nu}$ and any $\epsilon' > 0$, there exists some $\epsilon > 0$ such that
+The above result becomes obviously applicable thanks to the following expression for $T_2^2$.
+
+\begin{lemma}
+	\label{lemma: T2_analytic_form}
 	\begin{equation*}
-	\mathcal{W}(\mu, \nu ) < \epsilon \implies \sup_{t} \abs{F_{\mu}(t) - F_{\nu}(t)} \leq \epsilon'.
+	T_2^2 = a^T \Linv a  
 	\end{equation*}
@@
 
-\section{Proofs}
+\paragraph{Asymptotic null distribution for $T_2$.}
 
-\begin{proof}[Proof of Proposition \ref{prop: expectation_of_TG}]
-	Throughout, we will use the fact that $a_i | Z_i \sim \text{Rademacher}(\frac{p(Z_i)}{p(Z_i) + q(Z_i)})$, which is easily seen by an application of Bayes rule.
-
-	Begin by rewriting
+\begin{theorem}
+	\label{thm: asymptotic_null_distribution}
+	Denote the scaled version of the Laplacian smooth test statistic
 	\begin{equation*}
-	a^T L a = (\mathbf{h_0} + a - \mathbf{h_0})^T L (\mathbf{h_0} + a - \mathbf{h_0}) := (\mathbf{h_0} + \mathbf{\epsilon})^T L (\mathbf{h_0} + \mathbf{\epsilon}).
-	\end{equation*}
-	Expanding the quadratic form yields
-	\begin{equation*}
-	a^T L a = \mathbf{h_0}^T L \mathbf{h_0} + \mathbf{\epsilon}^T L \mathbf{\epsilon} + 2 \mathbf{h_0}^T L \mathbf{\epsilon}.
+	W_n = \sqrt{\frac{2}{\tr((\Linv)^2)}}\Bigl(T_2^2 - 4\tr(\Linv) \Bigr).
 	\end{equation*}
 
-	Going from back to front, we have that the first term has expectation $0$, because
+	If Conjectures \ref{conj: spectral_measure_conv} and \ref{conj: diagonal_entries} hold,
 	\begin{equation*}
-	\Expect{L_{ij} h_0(Z_i) \epsilon_{j}} = \Expect{L_{ij} h_0(Z_i) \Expect{\epsilon_{j} | Z} } = \Expect{L_{ij} h_0(Z_i) 0} = 0.
+	\lim_{n \to \infty} \sup_{t} \abs{\Prob{W_n \leq t} - \Phi(t)} = 0.
 	\end{equation*}
-
-	For the middle term, only the diagonal terms have non-zero expectation.
-	\begin{align*}
-	\Expect{ \mathbf{\epsilon}^T L \mathbf{\epsilon} } & = \sum_{i,j = 1}^{N} \Expect{L_{ij} \Expect{\epsilon_{j} \epsilon_i | Z}} \\
-	& \overset{(i)}{=} \sum_{1 \leq i < j \leq N} \Expect{L_{ij} \Expect{\epsilon_{j} | Z} \Expect{\epsilon_i | Z}} + \sum_{i = 1}^n \Expect{L_{ii}^2 \Expect{\epsilon_i^2 | Z}} \\
-	& = \sum_{i = 1}^N \Expect{L_{ii}^2 \Expect{\epsilon_i^2 | Z}}.
-	\end{align*}
-	where $(i)$ follows from the conditional independence relation $a_i \Perp a_j | Z$.
-
-	Then 
-	\begin{equation*}
-	\Expect{\epsilon_i^2 | Z} = \Expect{(a(Z_i) - h_0(Z_i))^2 | Z} = \Var{a(Z_i) | Z_i} = \frac{4}{N^2} \left(\frac{4 p(Z_i) q(Z_i)}{(p(Z_i) + q(Z_i))^2}\right)
-	\end{equation*}
-	and plugging this in, we have
-	\begin{align*}
-	\Expect{ \mathbf{\epsilon}^T L \mathbf{\epsilon} } & = \frac{16}{N^2}\sum_{i = 1}^{N} \Expect{L_{ii} \left(\frac{p(Z_i) q(Z_i)}{(p(Z_i) + q(Z_i))^2}\right)} \\
-	& = \frac{16}{N^2} \sum_{i = 1}^{N} \Expect{ \sum_{j \neq i}K(\norm{Z_i - Z_j}) \left(\frac{p(Z_i) q(Z_i)}{(p(Z_i) + q(Z_i))^2}\right)} \\
-	& = \frac{4 N}{N - 1} \int \int K(\norm{\x - \y}) \left[p(\x) q(\x)\right] \frac{p(\y) + q(\y)}{p(\x) + q(\x)} d\x d\y \\
-	\end{align*}
-
-	Using the relation $\frac{(a + b)^2 - (a - b)^2}{4} = ab$ yields the 1st and 2nd integrals of (\ref{eqn: expectation_of_TG}). The 3rd integral is exactly $\Expect{h_0^T L h_0}$.
-
-\end{proof}
-
-\begin{proof}[Proof of Proposition \ref{prop: expectation_of_TG_with_kde}]
-	Write $\mathbf{h} = \frac{\x - \y}{t}$. Via Taylor expansion, we can write
-	\begin{align*}
-	& \int K\left(\frac{\norm{\x - \y}}{t}\right) (p(\y) + q(\y) ) d\y \\
-	& \overset{(i)}{=} \int K(\norm{\mathbf{h}}) (p(\x) + q(\x) + \mathcal{O}(t \norm{h})) t^d d \mathbf{h} \\
-	& \overset{(ii)}{=} (p(\x) + q(\x)) + \mathcal{O}(t^{d + 1})
-	\end{align*}
-	where $(i)$ follows from the Lipschitz continuity of $p$ and $q$, and $(ii)$ follows from the integrability condition on $K$.
-
-	Applying this to the 2nd and 3rd integrals of (\ref{eqn: expectation_of_TG}) yields the two integrals of (\ref{eqn: expectation_of_TG_with_kde}).
-The 3rd integral is $\mathcal{O}(t^{d + 1})$ by Lemma \ref{prop: bousquet04}.
-\end{proof}
+\end{theorem}
 
 \begin{proof}
-	First, we rewrite $T_2$, using the fact that $a = \frac{2}{N} \left(\sum_{i \in \Lx} e_i - \sum_{i \in \Ly} e_i \right)$. 
-	\begin{align*}
-	a^T \Linv a & = \frac{4}{N^2} \left(\sum_{i, j \in \Lx }e_i \Linv e_j + \sum_{i, j \in \Ly }e_i \Linv e_j - 2\sum_{i \in \Lx, j \in \Ly} e_i \Linv e_j \right)\\
-	\end{align*}
-	Via this expression, we see that in the above summations
-	\begin{itemize}
-		\item For $i = j$, $e_i^T \Linv e_i$ appears exactly once.
-		\item For $i \neq  j$ and $i, j \in \Lx$ or $i,j \in \Ly$, $e_i^T \Linv e_j$ appears exactly twice.
-		\item For $i \in \Lx$, $j \in \Ly$, $- e_i^T \Linv e_j$ appears exactly twice.
-	\end{itemize}
-
-	Now, consider the expression
-	\begin{equation*}
-	 \sum_{u \in \Lx, v \in \Ly} R_{uv} - \sum_{u < v \in \Ly }R_{uv} - \sum_{u < v \in \Lx }R_{uv}.
-	\end{equation*}
-	Going from bottom to top, we have
-	\begin{itemize}
-		\item When $i \in \Lx$ and $j \in \Ly$, $R_{ij}$ will contribute $-2 e_i \Linv e_j$. No other $R_{uv}$ will contribute anything to this term.
-		\item When $i < j \in \Lx$ or $i < j \in \Ly$, the term $R_{ij}$ in the 2nd or 3rd sum will appear exactly once and will contribute $2 e_i \Linv e_j$.
-No other $R_{uv}$ will contribute anything to this term.
-		\item When $i = j \in \Lx$, $- R_{ik}$ will contribute $ - e_i \Linv e_i$ for each $k \neq i \in \Lx$, and will contribute $ e_i \Linv e_i$ for each $k \in \Ly$.
-The total contribution will be $ \left(\abs{\Ly} - \abs{\Lx} + 1\right) (e_i \Linv e_i) = e_i \Linv e_i$.
-The same reasoning holds for $i = j \in \Ly$.
-	\end{itemize}
-
-	All contributions from all $R_{uv}$ can be put into one of the three proceeding categories. Therefore,
-	\begin{equation*}
-	a^T \Linv a = \frac{4}{N^2} \left(\sum_{u \in \Lx, v \in \Ly} R_{uv} - \sum_{u < v \in \Ly }R_{uv} - \sum_{u < v \in \Lx }R_{uv}\right)
-	\end{equation*}
-
-	(\ref{eqn: expectation_of_T2}) follows from taking expectation and noting that $X_i$ and $X_j$ are identically distributed for all $i$ and $j$.
-
-\end{proof}
-
-\begin{proof}[Proof of Theorem \ref{thm: asymptotic_null_distribution}]
 	We will proceed by
@@
 \end{proof}
+\paragraph{Asymptotic mean.}
 
-\begin{proof}[Proof of Lemma \ref{lem: cond_expectation_of_T2}]
+Under the null hypothesis of the \textbf{related generative model}, we have
 \begin{align}
-	\Expect{T^2 \vert Z} & = \Expect{a^T \Linv a \vert Z} \nonumber \\
-	& \overset{(i)}{=} \sum_{i = 1}^{N} \sum_{j = 1}^{N} \Expect{a_i a_j} \Linv_{ij} \nonumber \\
-	& =  \sum_{i = 1}^{N} \frac{1}{4N^2} \Linv_{ii} \nonumber \\
-	& = \frac{\tr(\Linv)}{4N^2}.
+\Expect{T^2} & = \Expect{a^T \Linv a} \nonumber \\
+& = \Expect{ \Expect{a^T \Linv a \mid  Z_1^{n+m}} } \nonumber \\
+& \overset{(i)}{=} \Expect{ \sum_{i = 1}^{n + m} \sum_{j = 1}^{n + m} \Expect{a_i a_j} \Linv_{ij} } \nonumber \\
+& =  \Expect{\sum_{i = 1}^{n + m} \frac{1}{p(1 - p)} \Linv_{ii}} \nonumber \\
+& = \frac{1}{N^2 p(1 - p)}\cdot \Expect{\sum_{i = 1}^{n + m} \Linv_{ii} } \nonumber \\
+& = \frac{1}{N^2 p(1 - p)} \cdot \Expect{\tr(\Linv)} \label{eqn: asymptotic_expectation}
 \end{align}
 where $(i)$ comes from the independence of $Z$ and $a$ under $H_0$.
-\end{proof}
 
-\begin{proof}[Proof of Lemma \ref{lem: cond_var_of_T2}]
-	First, we re-arrange $T_2$. 
+\paragraph{Asymptotic variance.}
+
+We begin by computing $\Expect{(T^2)^2}$. We will need the following terms
 \begin{align*}
-	T_2 & = \sum_{i = 1}^{N} \sum_{j = 1}^{N} a_i a_j \Linv_{ij} \\
-	& = 2 \sum_{i \leq j} a_i a_j \Linv_{ij} - \frac{4}{N^2}\sum_{i = 1}^{N} \Linv_{ii}.
+S_4 & := \sum_{i, k} \Linv_{ii} \Linv_{kk} \\
+S_5 & := \sum_{i, j} \Linv_{ij} \Linv_{ij} \\
+S_6 & := \sum_{i} \Linv_{ii} \Linv_{ii}.
 \end{align*}
-	Therefore, for $R_i \overset{i.i.d}{\sim} \text{Rademacher}(1/2)$,
+Then
 \begin{align*}
-	\Var{T_2 \vert Z} & = 4 \Var{\sum_{i \leq j} a_i a_j \Linv_{ij} | Z} \\
-	& = \frac{64}{N^4} \Var{\sum_{i \leq j} R_i R_j \Linv_{ij} \vert Z } \\
-	& = \frac{32}{N^4} \tr[(\Linv)^2].
+\Expect{(T^2)^2} & = \Expect{\sum_{i,j,k,l} \Linv_{ij} \Linv_{kl} \Expect{a_i a_j a_k a_l} } \\
+& = \Expect{\frac{1}{N^2 p^2(1 - p)^2}(S_4 + 2 S_5 - 3 S_6) + \frac{p^3 + (1 - p)^3}{N^4 p^3(1 - p)^3}S_6 } \\
+& \overset{(i)}{=} \frac{1}{N^4 p^2(1 - p)^2} \left( \Expect{tr(\Linv)}^2 + 2 \Expect {tr(\Linv \Linv)} - 3 \Expect{S_6} \right) + \frac{p^3 + (1 - p)^3}{N^4 p^3(1 - p)^3} \Expect{S_6}
+\end{align*}
+where $(i)$ follows from Lemma \ref{lem: cross_moments}. Along with (\ref{eqn: asymptotic_expectation}) we therefore obtain
+\begin{equation*}
+\Var{T^2} = \frac{1}{N^4 p^2(1 - p)^2} \left( 2 \Expect {tr(\Linv \Linv)} - 3 \Expect{S_6} \right) + \frac{p^3 + (1 - p)^3}{N^4 p^3(1 - p)^3} \Expect{S_6}
+\end{equation*}
+
+\begin{lemma}
+	\label{lem: cross_moments}
+	\begin{align*}
+	S_4 & = tr(\Linv)^2 \\
+	S_5 & = tr(\Linv \Linv)
+	\end{align*}
+\end{lemma}
+\begin{proof}
+	\begin{align*}
+	S_4 & = \sum_{i, k} \Linv_{ii} \Linv_{kk} \\
+	& = \left(\sum_{k} \Linv_{kk} \right)  \left(\sum_{i} \Linv_{ii} \right) \\
+	& = tr(\Linv)^2.
+	\end{align*}
+
+	\begin{align*}
+	S_5  & := \sum_{i, j} \Linv_{ij} \Linv_{ij} \\
+	& = \norm{\Linv}_F = \tr(\Linv \Linv). 
 	\end{align*}
@@
 
+\begin{lemma}
+	\label{lem: diagonal_elements}
+
+	Write the Laplacian in terms of its spectral decomposition
+	\begin{equation*}
+	L = \sum_{k = 2}^{n} \lambda_k u_k u_k^T 
+	\end{equation*}
+	where $\lambda_1 = 0$. Then, if $\max_{k, i} u_k[i] \leq \frac{C}{\sqrt{N}}$,
+	we have
+	\begin{equation*}
+	S_6 \leq C^4 \frac{tr(\Linv)^2}{N}.
+	\end{equation*}
+\end{lemma}
+
+\begin{proof}
+	\begin{align*}
+	\Linv_{ii} & = \sum_{k = 1}^{N} \lambda_k (u_k[i])^2 \\
+	& \leq \sum_{k = 1}^{N} \lambda_k \frac{C^2}{N} \\
+	& = C^2 \frac{tr(\Linv)}{N}.
+	\end{align*}
+	Therefore
+	\begin{align*}
+	S_6 & = \sum_{i} \Linv_{ii} \Linv_{ii} \\
+	& \leq \sum_{i} C^4 \frac{tr(\Linv)^2}{N^2} \\
+	& = C^4 \frac{tr(\Linv)^2}{N}.
+	\end{align*}
+\end{proof}
+
+\begin{theorem}
+	\label{theorem: asymptotic_null_dist}
+	Assume Conjectures \ref{conj: spectral_measure_conv}, \ref{conj: max_entry}, and \ref{conj: diagonal_max_bound} hold. Then
+	\begin{equation*}
+	\rho_n n^{3/2} (T_2^2) \overset{L}{\to} N(0, \frac{2}{p^2(1-p)^2}\mathbb{E}_{\nu_{\infty}}{V^2})  
+	\end{equation*}
+\end{theorem}
+
+\begin{proof}
+	Denote $W(n) = \rho_n n^{3/2} (T_2^2)$ and note that
+	\begin{align*}
+	W(n) & = \frac{\rho_n}{\sqrt{n}} \left(\sum_{i = 1}^{N} \sum_{j = 1}^{N} \Linv_{ij} a_ia_j \right).
+	\end{align*}
+	By the independence of $a_i$ and $a_j$, $\Expect{a_ia_j \mid a_j} = 0$ when $i \neq j$. For the diagonal terms, we have that
+	\begin{align*}
+	\Linv_{ii} \frac{\rho_n}{\sqrt{n}} & \overset{(i)}{=} \mathcal{O}(\frac{tr(\Linv) \rho_n}{n} \frac{1}{\sqrt{n}}) \\
+	& \overset{n \to \infty}{=} O(\frac{1}{\sqrt{n}}) \mathbb{E}_{\nu_\infty}(V) = 0
+	\end{align*}
+	where $(i)$ follows from Conjecture \ref{conj: max_entry} and Lemma \ref{lem: diagonal_elements}. So, asymptotically, the conditional expectation $\Expect{\Linv_{ij} a_i a_j \mid a_j} = 0 $ a.s.
+
+	The variance of $W(n)$ is calculated as
+	\begin{align*}
+	Var(W(n)) & = \frac{N^4 \rho_n^2}{N} Var(T_2^2) \\
+	& \overset{(i)}{=} \frac{\rho_n^2}{N p^2(1 - p)^2} \left( 2 \Expect {tr(\Linv \Linv)} - 3 \Expect{S_6} \right) + \frac{\rho_n \left(p^3 + (1 - p)^3 \right)}{N p^3(1 - p)^3} \Expect{S_6} \\
+	& \overset{(ii)}{=} \frac{1}{p^2(1 - p)^2 } (2 \mathbb{E}_{v_{\infty}}(V^2) ) + \mathcal{O}(\max_i(\Linv_{ii})^2 \rho_n^2) \\
+	& \overset{(iii)}{ = } \frac{1}{p^2(1 - p)^2 } (2 \mathbb{E}_{v_{\infty}}(V^2) ).
+	\end{align*}
+	where $(i)$ follows from our calculation for asymptotic variance, $(ii)$ follows from Conjecture \ref{conj: spectral_measure_conv} and the definition of $S_6$ and $(iii)$ from Conjecture \ref{conj: diagonal_max_bound}.
+
+	We compute
+	\begin{align*}
+	\lim_{n \to \infty} \frac{\rho_n^4}{n^2}tr((\Linv)^4) = \lim_{n \to \infty} \frac{1}{n}\mathbb{E}_{\nu_{\infty}}(V^4) = 0.
+    \end{align*}
+	and roughly this should imply the fourth moment is asymptotically $0$.
+
+	Finally, we have
+	\begin{align*}
+	\sum_{j = 1}^n \sigma_{ij}^2 & = \sum_{j = 1}^{n} \frac{\rho_n^2 (\Linv_{ij})^2}{n p^2 (1 - p)^2} \\
+	 & = \frac{\rho_n^2 }{n p^2 (1 - p)^2} (\Linv_i)^T (\Linv_i) \\
+	 & = \frac{\rho_n^2 }{n p^2 (1 - p)^2} (\Linv \Linv)_{ii} \\
+	 & \overset{(i)}{\leq} \frac{\rho_n^2 }{n p^2 (1 - p)^2} C^2 \frac{tr(\Linv \Linv)}{N} \\
+	 & = \mathcal{O}(1/n).
+	\end{align*}
+
+	By Theorem 2.1 of (de Jong 87), we therefore have
+	\begin{equation*}
+	\rho_n n^{3/2} (T_2^2) \overset{L}{\to} N(0, \frac{2}{p^2(1-p)^2}\mathbb{E}_{\nu_{\infty}}{V^2}).
+	\end{equation*}
+\end{proof}
+
+\section{Proofs}
+\begin{proof}[Proof of Lemma \ref{lemma: T2_analytic_form}]
+	Take the Lagrangian
+	\begin{equation*}
+	L(\theta, \lambda) = - a^T \theta + \lambda \theta^T L \theta
+	\end{equation*}
+	and let
+	\begin{align*}
+	\lambda^{\star} & = \sqrt{a^T \Linv a} \\
+	\theta^{\star} & = \frac{a^T \Linv}{\lambda^{\star}}
+	\end{align*}
+
+	The KKT conditions tell us that if
+	\begin{align}
+	\frac{\partial L}{\partial \theta}(\lambda^{\star}, \theta^{\star}) & = 0 \label{eqn: stationarity}\\
+	{\theta}^{\star^{T}} L \theta^{\star} & = 1 \label{eqn : complementary_slackness} \\
+	\lambda^{\star} & \geq 0 \label{eqn: dual_feasibility}
+	\end{align}
+	then $\theta^{\star}$ is a primal solution.
+
+	 We can write
+	\begin{equation*}
+	\frac{\partial L}{\partial \theta} = -a^T + \lambda \theta^T L
+	\end{equation*}
+	and plugging in our choice for $\theta^{\star}$ yields
+	\begin{align*}
+	\frac{\partial L}{\partial \theta}(\lambda^{\star}, \theta^{\star}) & = -a^T + \lambda^{\star} a^T \Linv L \\
+	& \overset{(i)}{=} -a^T + a^T P_{1^{\perp}} \\
+	& \overset{(ii)}{=} -a^T + a^T = 0.
+	\end{align*}
+	where $(i)$ follows from the already stated fact that $L$ has one $0$ eigenvalue with constant eigenvector, and $(ii)$ from the fact that $a \perp {\bf 1}$.
+As a result, $(\lambda^{\star}, \theta^{\star})$ satisfy (\ref{eqn: stationarity}).
+
+	Then, we have
+	\begin{align*}
+	{\theta}^{\star^{T}} L \theta^{\star} & = \frac{a^T \Linv L \Linv a}{\lambda^{\star^2}} \\
+	& = \frac{a^T \Linv a}{\lambda^{\star^2}} = 1.
+	\end{align*}
+	satisfying (\ref{eqn : complementary_slackness}).
+
+	Finally, because $L^{\dagger}$ is a positive-definite matrix, for any vector $v ~ v^T\Linv v > 0$, and therefore $\lambda^{\star} \geq 0$, verifying (\ref{eqn: dual_feasibility}).
+
+	As a result, $\theta^{\star}$ minimizes $a^T \theta$ subject to the given constraint. Plugging in our expression for $\theta^{\star}$ yields
+	\begin{equation*}
+	T = a^T \theta^{\star} = \frac{a^T \Linv a}{\lambda^{\star}} = \sqrt{a^T \Linv a}.
+	\end{equation*}	
+\end{proof}
+
 \end{document}

### 2019-01-30 notes/2018.12.11.NotesForWeek.tex
added backlog of notes

--- 2018-12-19 notes/graphtestingmasternotes.tex
@@
 
-\title{Graph Testing Notes}
+\title{Graph Testing: Notes for the Week of 12/11 - 12/18}
 \author{Alden Green}
@@
 
-We are given two distributions, $P$ and $Q$, with the ability to sample from either one. Our goal is to test the hypothesis $H_0: P = Q$ vs. the alternative $H_a: P \neq Q$.
+We are given two distributions, $P$ and $Q$, defined over compact set $\X \subset \Rd$, with the ability to sample from either one.
+Our goal is to test the hypothesis $H_0: P = Q$ vs.
+the alternative $H_a: P \neq Q$.
 
@@
 Take $L = D - A$ to be the (unnormalized) \textbf{Laplacian matrix} of $A$ (where $D$ is the diagonal degree matrix with $D_{ii} = \sum_{j \in [n + m]} A_{ij}$).
+Let $M$ be the number of non-zero entries of $A$.
 Denote by $B$ the $M \times N$ \textbf{incidence matrix} of $A$, where we denote the $i$th row of $B$ as $B_i$ and set $B_i$ to have entry $A_{ij}$ in position $i$, $-A_{ij}$ in position $j$, and $0$ everywhere else.
-
-\paragraph{Resistance distances.}
-
-There are many distances one can define over nodes in a graph. The \textbf{resistance distance between nodes $u$ and $v$}, $R_{uv}$, is defined as
-\begin{equation*}
-R_{uv} = (e_u - e_v)^T \Linv (e_u - e_v).
-\end{equation*}
 
@@
 
-We begin by defining our \textbf{laplacian smooth} test statistic. 
+We define our \textbf{laplacian smooth} test statistic. 
 \begin{equation*}
@@
 
-(Bhattacharya 2018) defines a general notion of 2-sample \textbf{graph-based test statistics}
-\begin{equation*}
-T_{\mathcal{G}} = \frac{1}{N^2}\sum_{i = 1}^{n} \sum_{j = n + 1}^{n + m} A_{ij}
-\end{equation*}
-Although he develops theory for this statistic in the context of $k$NN and minimum spanning tree graphs, we will at present consider it for the complete weighted similarity graph defined by $A$ above.
-Then, we can write
-\begin{equation*}
-T_{\mathcal{G}} = a^T L a.
-\end{equation*}
-
-Finally, define $\mathcal{H}$ to be a \textbf{reproducing kernel Hilbert space} with $K$ the associated kernel.
-Let $\mathcal{F}$ be the unit ball of $\mathcal{H}$, and let the evaluation of $f \in \mathcal{F}$ at the sample points $Z_1, \ldots, Z_N$ be denoted by $\textbf{f} = f(Z_1, \ldots, Z_N)$.
-Then, the statistic $\text{MMD}_b$ of (Gretton 2012) can be written as
-\begin{equation*}
-T_{\mathcal{K}} = \sup_{f \in \mathcal{F}} a^T \textbf{f} = a^T K a. 
-\end{equation*}
-
 \paragraph{Distances between probability measures.}
-
-We will need distances between probability measures for two different purposes.
-The first is that they are self-evidently useful in analyzing limiting distributions of statistics (in particular in this case, our test statistics).
-
-For a function $f$, define its \textbf{Lipschitz norm} $\norm{f}_{L}$ to be
-\begin{equation*}
-\inf \set{K : \abs{f(x) - f(y)} \leq K \norm{x - y}}.
-\end{equation*}
-Define the \textbf{Wasserstein distance} between two measures $\mu$ and $\nu$ to be
-\begin{equation*}
-\mathcal{W}(\mu, \nu) := \sup \left \{ \abs{\int h \, d\mu - \int h \, d\nu}: h \text{ Lipschitz, with } \norm{h}_{L} \leq 1 \right\}.
-\end{equation*}
-
-If the measures $\mu$ and $\nu$ have corresponding cumulative distribution functions $F_{\mu}$ and $F_{\nu}$ then we can define the \textbf{Kolmogorov-Smirnov distance} to be
-\begin{equation*}
-\norm{F_\mu - F_\nu}_{\infty} := \sup_{t} \abs{F_{\mu}(t) - F_{\nu}(t)}.
-\end{equation*}
-
-The second reason we will use distances between probability measures is that they themselves make for good test statistics!
 
@@
 
-\paragraph{Holder functions.}
 We will show that the Laplacian constraint $\norm{B \theta}_2 \leq 1$ is very similar to the constraint $f_{\theta} \in B_W(X,\mu^2)$ for the right choice of $K$, over all Holder functions.
 
+\paragraph{Holder functions}
 For mapping $f: \Rd \to \Reals$ and $\beta$ a positive integer, we say $f$ is a \textbf{$\beta$-Holder function} if there exists $C > 0$ such that for all $x, y \in \X$
@@
 Roughly speaking, this means the functions have bounded $\beta$ partial derivatives.
-
-\section{Conjectures}
-
-Conjectures \ref{conj: spectral_measure_conv} and \ref{conj: diagonal_entries} will be needed for Theorem \ref{thm: asymptotic_null_distribution}.
-
-\begin{conjecture}
-	\label{conj: spectral_measure_conv}
-	There exists a sequence of scaling factors $(\rho_n)_{n = 1}^{\infty}$ such that the \textbf{spectral measure} $\mu_n$ of $\rho_n \Linv$ converges weakly in probability
-	\begin{equation*}
-	\mu_n(\rho_n \Linv) \overset{\ast}{\rightharpoonup} \nu_{\infty}.
-	\end{equation*}
-
-	where $V \sim \nu_{\infty}$ and $V_n \sim \mu_n$ are bounded almost surely for all $n$ by some constant $C$. 
-\end{conjecture}
-
-\begin{conjecture}
-	\label{conj: diagonal_entries}
-	For all $\epsilon > 0$, there exists $N$ such that
-	\begin{equation*}
-	\Prob{\underset{i \in [n]}{\max} \, \frac{1}{n} \left(\{ \rho_n \Linv\}^2 \right)_{ii} \leq \epsilon} \geq 1 - \epsilon
-	\end{equation*}
-	for all $n \geq N$.
-\end{conjecture}
 
@@
 	with high probability.
+
 \end{proof}
 
-\section{Results}
-
-\paragraph{Expectation of two-sample test statistics.}
-
-The expectation of the above statistics is potentially a good way to understand their large sample behavior, as quadratic forms often satisfy laws of large numbers assuming the matrices are well-conditioned.
-
-\begin{proposition}
-	\label{prop: expectation_of_TG}
-	Draw $Z$ and $a$ under the binomial data model, and assume both $P$ and $Q$ are absolutely continuous with respect to Lebesgue measure over Euclidean space $\Reals^d$.
-Write $h_0(x) = p(x) - q(x)$ with empirical analogue $\mathbf{h_0} = (h_0(Z_1), \ldots, h_0(Z_n) )$.
-Then
-	\begin{align}
-	\label{eqn: expectation_of_TG}
-	\Expect{T_{\mathcal{G}}} = & \int \int K(\norm{\x - \y}) \left[p(\x) + q(\x)\right] \left[p(\y) + q(\y)\right] d\x d\y \nonumber \\
-	- & \frac{N}{N - 1} \int \int K(\norm{\x - \y}) \left[h_0(\x)\right]^2 \frac{p(\y) + q(\y)}{p(\x) + q(\x)} d\x d\y \nonumber \\
-	+ & \frac{N}{N - 1} \int \int K(\norm{\x - \y}) \left[h_0(\x) - h_0(\y)\right]^2 \left[p(\y) + q(\y)\right] \left[p(\x) + q(\x)\right] d\x d\y.
-	\end{align}
-\end{proposition}
-
-Note that even under the null hypothesis, where $h_0 = 0$,
-\begin{equation*}
-\Expect{T_{\mathcal{G}}} = \int \int K(\norm{\x - \y}) \left[p(\x) + q(\x)\right] \left[p(\y) + q(\y)\right] d\x d\y
-\end{equation*}
-which is not distribution-free, unlike in the case of the $k$NN graph.
-
-Another interesting consequence of Proposition \ref{prop: expectation_of_TG} comes when we take $K(\x,\y) = K(\frac{\norm{\x - \y}}{t})$ and let $t \to 0$.
-
-\begin{proposition}
-	\label{prop: expectation_of_TG_with_kde}
-	If $p$ and $q$ are Lipschitz continuous functions with bounded hessians, and $K$ is a continuous function on $R^+$ such that $x^{2 + d} K(x) \in L_2$, then under the same setup as in Proposition \ref{prop: expectation_of_TG},
-	\begin{equation}
-	\label{eqn: expectation_of_TG_with_kde}
-	\frac{N - 1}{N}\underset{t \to 0}{\lim} \frac{1}{t^d} \Expect{T_{\mathcal{G}}} = \int  h_0(\x)^2 d\x + \int (p(\x) + q(\x))^2 d\x
-	\end{equation}
-\end{proposition}
-
-We turn now to the expectation of the Laplacian smooth statistic.
-\begin{proposition}
-	\label{prop: expectation_of_T2}
-	Under the fixed label data model
-	\begin{equation}
-	\label{eqn: expectation_of_T2}
-	\Expect{a^T \Linv a} = \Expect{R_{X_1 Y_1}} - \frac{N - 1}{2N} \Expect{R_{X_1,X_2}} - \frac{N - 1}{2N} \Expect{R_{Y_1, Y_2}}
-	\end{equation}
-\end{proposition}
-
-Note that, although the resistance distances are between only two nodes, in each case the expectation is over the entire (random) graph $G$.
-
-\paragraph{Asymptotic null distribution for $T_2$.}
-
-We can compute an asymptotic null distribution for $T_2$, although its formulation depends on the eigenvalues of the matrix $\Linv$ which themselves are not obvious.
-
-\begin{theorem}
-	\label{thm: asymptotic_null_distribution}
-	Denote the scaled version of the Laplacian smooth test statistic
-	\begin{equation*}
-	W_n = \sqrt{\frac{N^4}{32 \cdot \tr((\Linv)^2)}}\Bigl(T_2^2 - \frac{\tr(\Linv)}{4N^2} \Bigr).
-	\end{equation*}
-
-	If Conjectures \ref{conj: spectral_measure_conv} and \ref{conj: diagonal_entries} hold,
-	\begin{equation*}
-	\lim_{n \to \infty} \sup_{t} \abs{\Prob{W_n \leq t} - \Phi(t)} = 0.
-	\end{equation*}
-\end{theorem}
-
-To prove Theorem \ref{thm: asymptotic_null_distribution}, we will need the following calculations of moments under $H_0$.
-
-\begin{lemma}
-	\label{lem: cond_expectation_of_T2}
-	Under $H_0$, the conditional expectation $\Expect{T_2 \vert Z} = \frac{\tr(\Linv)}{N^2}$. 
-\end{lemma}
-
-\begin{lemma}
-	\label{lem: cond_var_of_T2}
-	Under $H_0$, the conditional variance $\Var{T_2 \vert Z} = \frac{32 \tr[(\Linv)^2]}{N^4}$. 
-\end{lemma}
-
-\section{Supplemental Results}
+\section{SUPPLEMENTAL RESULTS}
 
@@
 
-\paragraph{Regularization functionals.}
-
-When taking the supremum over functions which satisfy $\norm{B \theta}_2 \leq 1$, we will argue that this constraint is well-behaved in the limit, i.e.
-that it converges to the \textbf{regularization functional} $\norm{\cdot}_{W_0^{1,2}(\X, \mu^2)}$.
-Proposition \ref{prop: convprob_regularization_functional} makes this convergence uniform over the set of $3$-Holder functions (essentially functions with bounded $3rd$ derivative).
-Proposition \ref{prop: bousquet04} makes this convergence only pointwise, but merely requires that $f$ have bounded 2nd derivative.
+\paragraph{Regularization functional.}
 
@@
 
-\begin{proposition}[Bousquet 04]
-	\label{prop: bousquet04}
-	If $p$ and $q$ are Lipschitz continuous functions with bounded hessians, and $K$ is a continuous function on $R^+$ such that $x^{2 + d} K(x) \in L_2$, then 
-	\begin{align}
-	\label{eqn: bousquet04}
-	& \underset{t \to 0}{\lim} \frac{d}{Ct^{d+2}} \int  K(\norm{\x - \y}/ t)(h_0(\x) - h_0(\y))^2 (p(\x) + q(\x))(p(\y) + q(\y)) d\x d \y \nonumber \\
-	& = \int \norm{\nabla h_0(\x)}^2 (p(\x) + q(\x))^2 d\x
-	\end{align}
-\end{proposition}
-
-\begin{lemma}[von Luxburg 12]
-	\label{lem: vonluxburg12}
-	Assume $P$ and $Q$ are absolutely continuous with respect to Lebesgue measure on Euclidean space $\Reals^d$, with density functions $p$ and $q$, respectively.
-Let $K(x,y) = \frac{1}{(2\pi \sigma^2)}^{d/2} \exp{-\frac{\norm{x - y}^2}{2 \sigma^2}}$.
-
-	Under some regularity assumptions on $p$ and $q$, if $n \to \infty$, $\sigma \to 0$ , and $n\sigma^{d+2}/ \log(n) \to \infty$, then
-	\begin{align*}
-	n R_{XY} \to \frac{2}{p(X) + q(X)} + \frac{2}{p(Y) + q(Y)} \textit{ almost surely }
-	\end{align*}
-	with equivalent statements holding for $X_1, X_2$ and $Y_1, Y_2$.  
-\end{lemma}
-
-\paragraph{Central limit theorem for quadratic forms.}
-
-\begin{theorem}[Chatterjee 08]
-	\label{thm: clt_quadratic_forms}
-	Let $a = (a_1, \ldots, a_n)$ be i.i.d random variables with with $\Prob{X_i = 1} = \Prob{X_i = -1} = 1/2$. For some fixed real valued symmetric matrix $M = (M_{ij})_{1 \leq i, j \leq n}$, define
-	\begin{equation*}
-	W = a^T M a.
-	\end{equation*}
-	with $\mu$ denoting the law of $(W - EW) / \sqrt{\Var{W}}$.
-
-	Then, letting $\mathcal{G}$ be the standard Gaussian measure 
-	\begin{equation}
-	\label{eqn: wass_distance_to_normal}
-	\mathcal{W}(\mu, \mathcal{G} ) \leq \left(\frac{\tr(M^4)}{\tr(M^2)^2} \right)^{1/2} + \left( \frac{5 \, \underset{i}{\max} \, (M_{ii})^2}{\tr(M^2)} \right)^{1/2}.
-	\end{equation}
-\end{theorem}
-
-\paragraph{Translating from Wasserstein to Kolmogorov distance.}
-
-\begin{lemma}[Wasserstein to Kolmogorov distance]
-	\label{lem: wass_to_ks_distance}
-	For any probability measures $\mu$, $\nu$ with corresponding cdfs $F_{\mu}$ and $F_{\nu}$ and any $\epsilon' > 0$, there exists some $\epsilon > 0$ such that
-	\begin{equation*}
-	\mathcal{W}(\mu, \nu ) < \epsilon \implies \sup_{t} \abs{F_{\mu}(t) - F_{\nu}(t)} \leq \epsilon'.
-	\end{equation*}
-\end{lemma}
-
-\section{Proofs}
-
-\begin{proof}[Proof of Proposition \ref{prop: expectation_of_TG}]
-	Throughout, we will use the fact that $a_i | Z_i \sim \text{Rademacher}(\frac{p(Z_i)}{p(Z_i) + q(Z_i)})$, which is easily seen by an application of Bayes rule.
-
-	Begin by rewriting
-	\begin{equation*}
-	a^T L a = (\mathbf{h_0} + a - \mathbf{h_0})^T L (\mathbf{h_0} + a - \mathbf{h_0}) := (\mathbf{h_0} + \mathbf{\epsilon})^T L (\mathbf{h_0} + \mathbf{\epsilon}).
-	\end{equation*}
-	Expanding the quadratic form yields
-	\begin{equation*}
-	a^T L a = \mathbf{h_0}^T L \mathbf{h_0} + \mathbf{\epsilon}^T L \mathbf{\epsilon} + 2 \mathbf{h_0}^T L \mathbf{\epsilon}.
-	\end{equation*}
-
-	Going from back to front, we have that the first term has expectation $0$, because
-	\begin{equation*}
-	\Expect{L_{ij} h_0(Z_i) \epsilon_{j}} = \Expect{L_{ij} h_0(Z_i) \Expect{\epsilon_{j} | Z} } = \Expect{L_{ij} h_0(Z_i) 0} = 0.
-	\end{equation*}
-
-	For the middle term, only the diagonal terms have non-zero expectation.
-	\begin{align*}
-	\Expect{ \mathbf{\epsilon}^T L \mathbf{\epsilon} } & = \sum_{i,j = 1}^{N} \Expect{L_{ij} \Expect{\epsilon_{j} \epsilon_i | Z}} \\
-	& \overset{(i)}{=} \sum_{1 \leq i < j \leq N} \Expect{L_{ij} \Expect{\epsilon_{j} | Z} \Expect{\epsilon_i | Z}} + \sum_{i = 1}^n \Expect{L_{ii}^2 \Expect{\epsilon_i^2 | Z}} \\
-	& = \sum_{i = 1}^N \Expect{L_{ii}^2 \Expect{\epsilon_i^2 | Z}}.
-	\end{align*}
-	where $(i)$ follows from the conditional independence relation $a_i \Perp a_j | Z$.
-
-	Then 
-	\begin{equation*}
-	\Expect{\epsilon_i^2 | Z} = \Expect{(a(Z_i) - h_0(Z_i))^2 | Z} = \Var{a(Z_i) | Z_i} = \frac{4}{N^2} \left(\frac{4 p(Z_i) q(Z_i)}{(p(Z_i) + q(Z_i))^2}\right)
-	\end{equation*}
-	and plugging this in, we have
-	\begin{align*}
-	\Expect{ \mathbf{\epsilon}^T L \mathbf{\epsilon} } & = \frac{16}{N^2}\sum_{i = 1}^{N} \Expect{L_{ii} \left(\frac{p(Z_i) q(Z_i)}{(p(Z_i) + q(Z_i))^2}\right)} \\
-	& = \frac{16}{N^2} \sum_{i = 1}^{N} \Expect{ \sum_{j \neq i}K(\norm{Z_i - Z_j}) \left(\frac{p(Z_i) q(Z_i)}{(p(Z_i) + q(Z_i))^2}\right)} \\
-	& = \frac{4 N}{N - 1} \int \int K(\norm{\x - \y}) \left[p(\x) q(\x)\right] \frac{p(\y) + q(\y)}{p(\x) + q(\x)} d\x d\y \\
-	\end{align*}
-
-	Using the relation $\frac{(a + b)^2 - (a - b)^2}{4} = ab$ yields the 1st and 2nd integrals of (\ref{eqn: expectation_of_TG}). The 3rd integral is exactly $\Expect{h_0^T L h_0}$.
-
-\end{proof}
-
-\begin{proof}[Proof of Proposition \ref{prop: expectation_of_TG_with_kde}]
-	Write $\mathbf{h} = \frac{\x - \y}{t}$. Via Taylor expansion, we can write
-	\begin{align*}
-	& \int K\left(\frac{\norm{\x - \y}}{t}\right) (p(\y) + q(\y) ) d\y \\
-	& \overset{(i)}{=} \int K(\norm{\mathbf{h}}) (p(\x) + q(\x) + \mathcal{O}(t \norm{h})) t^d d \mathbf{h} \\
-	& \overset{(ii)}{=} (p(\x) + q(\x)) + \mathcal{O}(t^{d + 1})
-	\end{align*}
-	where $(i)$ follows from the Lipschitz continuity of $p$ and $q$, and $(ii)$ follows from the integrability condition on $K$.
-
-	Applying this to the 2nd and 3rd integrals of (\ref{eqn: expectation_of_TG}) yields the two integrals of (\ref{eqn: expectation_of_TG_with_kde}).
-The 3rd integral is $\mathcal{O}(t^{d + 1})$ by Lemma \ref{prop: bousquet04}.
-\end{proof}
-
-\begin{proof}
-	First, we rewrite $T_2$, using the fact that $a = \frac{2}{N} \left(\sum_{i \in \Lx} e_i - \sum_{i \in \Ly} e_i \right)$. 
-	\begin{align*}
-	a^T \Linv a & = \frac{4}{N^2} \left(\sum_{i, j \in \Lx }e_i \Linv e_j + \sum_{i, j \in \Ly }e_i \Linv e_j - 2\sum_{i \in \Lx, j \in \Ly} e_i \Linv e_j \right)\\
-	\end{align*}
-	Via this expression, we see that in the above summations
-	\begin{itemize}
-		\item For $i = j$, $e_i^T \Linv e_i$ appears exactly once.
-		\item For $i \neq  j$ and $i, j \in \Lx$ or $i,j \in \Ly$, $e_i^T \Linv e_j$ appears exactly twice.
-		\item For $i \in \Lx$, $j \in \Ly$, $- e_i^T \Linv e_j$ appears exactly twice.
-	\end{itemize}
-
-	Now, consider the expression
-	\begin{equation*}
-	 \sum_{u \in \Lx, v \in \Ly} R_{uv} - \sum_{u < v \in \Ly }R_{uv} - \sum_{u < v \in \Lx }R_{uv}.
-	\end{equation*}
-	Going from bottom to top, we have
-	\begin{itemize}
-		\item When $i \in \Lx$ and $j \in \Ly$, $R_{ij}$ will contribute $-2 e_i \Linv e_j$. No other $R_{uv}$ will contribute anything to this term.
-		\item When $i < j \in \Lx$ or $i < j \in \Ly$, the term $R_{ij}$ in the 2nd or 3rd sum will appear exactly once and will contribute $2 e_i \Linv e_j$.
-No other $R_{uv}$ will contribute anything to this term.
-		\item When $i = j \in \Lx$, $- R_{ik}$ will contribute $ - e_i \Linv e_i$ for each $k \neq i \in \Lx$, and will contribute $ e_i \Linv e_i$ for each $k \in \Ly$.
-The total contribution will be $ \left(\abs{\Ly} - \abs{\Lx} + 1\right) (e_i \Linv e_i) = e_i \Linv e_i$.
-The same reasoning holds for $i = j \in \Ly$.
-	\end{itemize}
-
-	All contributions from all $R_{uv}$ can be put into one of the three proceeding categories. Therefore,
-	\begin{equation*}
-	a^T \Linv a = \frac{4}{N^2} \left(\sum_{u \in \Lx, v \in \Ly} R_{uv} - \sum_{u < v \in \Ly }R_{uv} - \sum_{u < v \in \Lx }R_{uv}\right)
-	\end{equation*}
-
-	(\ref{eqn: expectation_of_T2}) follows from taking expectation and noting that $X_i$ and $X_j$ are identically distributed for all $i$ and $j$.
-
-\end{proof}
-
-\begin{proof}[Proof of Theorem \ref{thm: asymptotic_null_distribution}]
-	We will proceed by
-	\begin{enumerate}
-		\item Conditioning on the high-probability outcome that the Laplacian converges to a limiting object in the right sense.
-		\item Showing that, under such convergence of the Laplacian, both terms in Theorem \ref{thm: clt_quadratic_forms} grow small with $n$.
-		\item Converting from Wasserstein distance to Kolmogorov distance. 
-	\end{enumerate}
-
-	\paragraph{Step 1.}
-	Fix $\epsilon > 0$. Throughout, let $P_Z$ denote the distribution of $Z$, and likewise $P_a$ denote the distribution of $a$.
-
-	For  $V_n \sim \nu_n(\rho_n \Linv)$, and $V \sim \nu_{\infty}$ let
-	\begin{equation*}
-	A_n = \biggl \{z \in \mathbb{R}^n: \abs{ E V_n^p - E V^p } \leq \epsilon \text{ for } p = 1, 2, 4 \biggr \} \bigcup \set{ z \in \mathbb{R}^n: \underset{i \in [n]}{\max} \, \frac{1}{n} \left(\{ \rho_n \Linv\}^2 \right)_{ii} \leq \epsilon}.
-	\end{equation*}
-
-	It is not hard to see that our Conjectures \ref{conj: spectral_measure_conv} and \ref{conj: diagonal_entries} imply $A_n$ will eventually have high probability.
-	\begin{align}
-	\label{eqn: good_event_probability}
-	\Prob{A_n} & \geq \Prob{\biggl \{z \in \mathbb{R}^n: \abs{ E V_n^p - E V^p } \leq \epsilon \biggr \}} + \Prob{\set{ z \in \mathbb{R}^n: \underset{i \in [n]}{\max} \, \frac{1}{n} \left(\{ \rho_n \Linv\}^2 \right)_{ii} \leq \epsilon}} \nonumber \\
-	& \overset{(i)}{\geq} 1 - 2 \epsilon \text{ for all } n \geq N.
-	\end{align}
-	where $(i)$ follows from Conjecture \ref{conj: diagonal_entries} (for the second term), and Conjecture \ref{conj: spectral_measure_conv} (for the first term).
-
-	Writing $W_n := W_n(z,a)$ to emphasize that it is a function of $z$ and $a$, we have by Tonelli's theorem that
-	\begin{align}
-	\label{eqn: conditional_ks_distance}
-	\sup_{t} \abs{\Prob{W_n \leq t} - \Phi(t)} & \overset{(i)}{=} \sup_{t} \abs{ \int_{\Reals^N} \left( \int_{\{-1,1\}^N} 1(W_n(z,a) \leq t) dP_a\right) dP_z - \Phi(t) } \nonumber \\
-	& = \sup_{t} \abs{ \int_{\Reals^N} \left( \int_{\{-1,1\}^N} 1(W_n(z,a) \leq t) dP_a\right) - \Phi(t)dP_z } \nonumber \\
-	& \leq \int_{\Reals^N} \sup_{t} \abs{\left( \int_{\{-1,1\}^N} 1(W_n(z,a) \leq t)   dP_a\right) - \Phi(t) } dP_z \nonumber \\
-	& \overset{(ii)}{\leq} \int_{A_n} \sup_{t} \abs{\left( \int_{\{-1,1\}^N} 1(W_n(z,a) \leq t) dP_a\right) - \Phi(t)} dP_z + 2 \epsilon
-	\end{align}
-	where $(i)$ follows from Tonelli's theorem and $(ii)$ from (\ref{eqn: good_event_probability}).
-
-	\paragraph{Step 2.}
-
-	Denote as
-	\begin{equation*}
-	F_{a|z}(z, t) := \left( \int_{\{-1,1\}^N} 1(W_n(z,a) \leq t) dP_a\right)
-	\end{equation*}
-	and note that for any $z$ this defines a measure over the Borel $\sigma$-algebra $\mathcal{B}(\Reals)$, which we will call $\mu_{a|Z}(z)$.
-
-	We wish to upper bound $\mathcal{W}(\mu_{a|Z}(z), \mathcal{G})$. To do so, we will compute upper bounds for each present in (\ref{eqn: wass_distance_to_normal}). For the first term, we have
-	\begin{align*}
-	\frac{\tr(\{\Linv \}^4)}{\tr(\{ \Linv \}^2)^2} & = \frac{1}{n} \frac{\frac{1}{n} \tr(\rho_n^4 \{\Linv \}^4)}{ \frac{1}{n^2} \rho_n^4 \tr(\{ \Linv \}^2)^2} \\
-	& \leq \frac{1}{n} \frac{\Expect{V^4} + \epsilon}{\Expect{V^2}^2 - \epsilon}.
-	\end{align*}
-	For the second term, we have
-	\begin{align*}
-	\frac{ \underset{i}{\max}(\{\Linv\}^2)_{ii} }{ \tr(\{\Linv\}^2) } & = \frac{ \frac{\rho_n^2}{n} (\{\Linv\}^2)_{ii}}{ \frac{\rho_n^2}{n} \tr(\{\Linv\}^2) } \\
-	& \leq \frac{ \epsilon }{ \Expect{V^2} - \epsilon }.
-	\end{align*}
-	By Theorem \ref{thm: clt_quadratic_forms} we therefore have
-	\begin{equation}
-	\label{eqn: wasserstein_conditional_bound}
-	\mathcal{W}(\mu_{a|Z}(z), \mathcal{G} ) \leq \frac{1}{n} \frac{\Expect{V^4} + \epsilon}{\Expect{V^2}^2 - \epsilon} + \left(\frac{ \epsilon }{ \Expect{V^2} - \epsilon } \right)^{1/2}.
-	\end{equation}
-
-	\paragraph{Step 3.}
-	Note that the right hand side of (\ref{eqn: wasserstein_conditional_bound}) converges to 0 with $\epsilon$.
-Therefore, for any $\epsilon$ sufficiently small, by (\ref{eqn: wasserstein_conditional_bound}) and Lemma \ref{lem: wass_to_ks_distance} we have
-	\begin{equation*}
-	\norm{F_{Z | a} - \Phi}_{\infty} \leq \epsilon'.
-	\end{equation*}
-	Combined with (\ref{eqn: conditional_ks_distance}) we have
-	\begin{equation*}
-	\sup_{t} \abs{\Prob(W_n \leq t) - \Phi(t)} \leq 2 \epsilon + \epsilon'.
-	\end{equation*}
-	for all $n \geq n_0$.
-
-\end{proof}
-
-\begin{proof}[Proof of Lemma \ref{lem: cond_expectation_of_T2}]
-	\begin{align}
-	\Expect{T^2 \vert Z} & = \Expect{a^T \Linv a \vert Z} \nonumber \\
-	& \overset{(i)}{=} \sum_{i = 1}^{N} \sum_{j = 1}^{N} \Expect{a_i a_j} \Linv_{ij} \nonumber \\
-	& =  \sum_{i = 1}^{N} \frac{1}{4N^2} \Linv_{ii} \nonumber \\
-	& = \frac{\tr(\Linv)}{4N^2}.
-	\end{align}
-	where $(i)$ comes from the independence of $Z$ and $a$ under $H_0$.
-\end{proof}
-
-\begin{proof}[Proof of Lemma \ref{lem: cond_var_of_T2}]
-	First, we re-arrange $T_2$. 
-	\begin{align*}
-	T_2 & = \sum_{i = 1}^{N} \sum_{j = 1}^{N} a_i a_j \Linv_{ij} \\
-	& = 2 \sum_{i \leq j} a_i a_j \Linv_{ij} - \frac{4}{N^2}\sum_{i = 1}^{N} \Linv_{ii}.
-	\end{align*}
-	Therefore, for $R_i \overset{i.i.d}{\sim} \text{Rademacher}(1/2)$,
-	\begin{align*}
-	\Var{T_2 \vert Z} & = 4 \Var{\sum_{i \leq j} a_i a_j \Linv_{ij} | Z} \\
-	& = \frac{64}{N^4} \Var{\sum_{i \leq j} R_i R_j \Linv_{ij} \vert Z } \\
-	& = \frac{32}{N^4} \tr[(\Linv)^2].
-	\end{align*}
-\end{proof}
+\section{PROOFS}
 

### 2019-01-30 notes/2018.12.4.NotesForWeek.tex
added backlog of notes

\documentclass{article}
\usepackage{amsmath, amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Var}[1]{\mathrm{Var}\left( #1 \right)}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\Perp}{\perp \! \! \! \perp}

\newcommand{\Linv}{L^{\dagger}}
\newcommand{\tr}{\text{tr}}
\newcommand{\h}{\textbf{h}}
% \newcommand{\l}{\ell}
\newcommand{\x}{\textbf{x}}
\newcommand{\y}{\textbf{y}}
\newcommand{\z}{\textbf{z}}
\newcommand{\Lx}{\mathcal{L}_X}
\newcommand{\Ly}{\mathcal{L}_Y}
\DeclareMathOperator*{\argmin}{argmin}

\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\R}{\Reals}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}
\newtheorem{remark}{Remark}

\begin{document}

\title{Notes for the week 12/4 - 12/10}
\author{Alden Green}
\date{\today}
\maketitle

\section{Setup}

\paragraph{Data model.}

We are given two distributions, $P$ and $Q$, with the ability to sample from either one. Our goal is to test the hypothesis $H_0: P = Q$ vs. the alternative $H_a: P \neq Q$.

Under the \textbf{binomial data model}, our sampling procedure is to draw i.i.d Rademacher labels $L_i \in \set{1, -1}$ for $i \in \set{1, \ldots, N}$, and then sample 
$Z_i \sim P$ if $L_i = 1$ and $Z_i \sim Q$ otherwise. Define $1_X$ to be the length-$N$ indicator vector for $L_i = 1$
\begin{equation*}
1_X[i] = 
\begin{cases}
1, L_i = 1\\
0 \text{ otherwise } 
\end{cases}
\end{equation*}
and similarly for $1_Y$
\begin{equation*}
1_Y[j] = 
\begin{cases}
1, L_i= -1 \\
0 \text{ otherwise } 
\end{cases}
\end{equation*}
and define $a = \frac{1_X}{N / 2} - \frac{1_Y}{N / 2}$.

Under the \textbf{fixed label data model} we use the same data generating process as above, except fix $\mathcal{L}_X = \set{1, \ldots, N/2}$ and $\mathcal{L}_Y = \set{N/2, \ldots, N}$. Say that $L_i = 1$ for $i \in \mathcal{L}_X$ and $L_i = -1$ for $i \in \mathcal{L}_Y$, and call $\set{ X_1, \ldots, X_{\abs{\Lx}} } = \set{Z_i: i \in \mathcal{L}_X}$ and likewise for $Y$.

\paragraph{Graph.}

Form an $N \times N$ Gram matrix $A$, where $A_{ij} = K(Z_i, Z_j)$ for \textbf{kernel function} $K: \mathcal{X} \times \mathcal{X} \to [0, \infty)$. Let $G = (V,E)$ with $V = \set{Z_1, \ldots, Z_n}$ and $E = \set{A_{ij}: 1 \leq i < j \leq n}$. Take $L = D - A$ to be the (unnormalized) \textbf{Laplacian matrix} of $A$ (where $D$ is the diagonal degree matrix with $D_{ii} = \sum_{j \in [n + m]} A_{ij}$). Denote by $B$ the $N \times N^2$ \textbf{incidence matrix} of $A$, where the $i$th column of $B = B_i$ has entry $A_{ij}$ in position $i$, $-A_{ij}$ in position $j$, and $0$ everywhere else.

\paragraph{Resistance distances.}

There are many distances one can define over nodes in a graph. The \textbf{resistance distance between nodes $u$ and $v$}, $R_{uv}$, is defined as
\begin{equation*}
R_{uv} = (e_u - e_v)^T \Linv (e_u - e_v).
\end{equation*}

\paragraph{Holder condition}

We say a function $f: \Rd \to \R$ is \textbf{$\alpha$-Holder continuous} when
\begin{equation*}
\abs{f(x) - f(y)} \leq \norm{x - y}^{\alpha}.
\end{equation*}

We will require this condition so that degrees in geometric graphs are well-behaved in the limit.

\section{Desiderata}

\begin{itemize}
	\item Let $K$ be a \textbf{uniform kernel of radius $\epsilon$}, meaning 
	\begin{equation*}
	K(x,y) = I(\norm{x - y} \leq \epsilon).
	\end{equation*}

	Assume $P$ and $Q$ have densities $p$ and $q$ with respect to Lebesgue measure. Say that for some $\alpha > 0$, $p$ and $q$ are $\alpha$-holder continuous. For the graph $G$ corresponding to the matrix $A$, with accompanying resistance distances, we wish to upper bound
	\begin{equation*}
	\abs{N \epsilon^d \Expect{R_{XY}} - \Expect{\frac{2}{p(X) + q(X)} + \frac{2}{p(Y) + q(Y)}}}
	\end{equation*}
\end{itemize}

\section{Supplemental Results}

Lemma \ref{lem: vonluxburg_poincare} follows from an application of a discrete version of Poincare's inequality. See (von Luxburg 12) for proof and details.
\begin{lemma}
	\label{lem: vonluxburg_poincare}
	For some $\widetilde{N}_{\max}, \widetilde{N}_{\min}, d_{\max}, d_{\min}$, for all $i \neq j$
	\begin{equation*}
	\abs{R_{ij} - \left(\frac{1}{d_i} + \frac{1}{d_j} \right)} \leq 2 a_1 \frac{1}{N \epsilon^{d+2}} \left( \frac{d_{\max}^2}{d_{\min}^3} \cdot(1 + 2 \frac{\widetilde{N}_{\max}^2}{\widetilde{N}_{\min}^2}) \right)
	\end{equation*}
	where $a_1 = \left( \frac{d \sqrt{d + 3}}{L_{\min}} \right)^{d + 1}$. 
\end{lemma}

Lemmas \ref{}

\begin{lemma}
Denote
\begin{alignat*}{2}
\mu_{\max} & := N \epsilon^d \nu_{d} (p_{\max} + q_{\max})/2,  ~~~ \mu_{\min} && := N \epsilon^d \nu_{d} (p_{\min} + q_{\min})/2 \beta
\end{alignat*}

and let $a_2 = \left(\frac{L_{\min}}{L_{\max}}\right)^d \frac{\nu_d}{2^d (d + 3)^{d/2}}$, $a_3 = \frac{\sqrt{d + 1}}{L_{\min}^d}$.

For $\widetilde{N}_{\max}, \widetilde{N}_{\min}, d_{\max}, d_{\min}$ as in Lemma \ref{lem: vonluxburg_poincare}, the following bounds hold
\begin{align*}
\Prob{\widetilde{N}_{\max} \geq (1 + z) \mu_{\max}} & \leq \frac{a_3}{\epsilon^d} \cdot \exp(- z^2 \mu_{\max} / 3) \\
\Prob{\widetilde{N}_{\min} \leq a_2 (1 - z) \mu_{\min}} & \leq \frac{a_3}{\epsilon^d} \cdot \exp(- z^2 a_2 \mu_{\min} / 3) \\
\Prob{d_{\max} \geq (1 + z) \mu_{\max}} & \leq n \cdot \exp(- z^2 \mu_{\max} / 3) \\
\Prob{d_{\min} \leq (1 - z) \mu_{\min}} & \leq n \cdot \exp(- z^2 \mu_{\min} / 3)
\end{align*}
\end{lemma}

\begin{lemma}
	\label{lem: prob_to_expectation}
	For random variable $X$ satisfying
	\begin{equation*}
	\Prob{X \leq (1 - z)\mu_n} \leq \exp(-z^2 \mu_n/3 + \log n)
	\end{equation*}
	the inverse moment $\Expect{\frac{1}{(1 + X)^k}}$, $k > 0$, satisfies for any $z < 1$ 
	\begin{equation*}
	\Expect{\frac{1}{(1 + X)^k}} \leq \exp(-z^2 \mu_n/3 + \log n) + \frac{1}{(1 + \mu_n(1 - z))^k}
	\end{equation*}

	Similarly, for random variable $Y$ satisfying
	\begin{equation*}
	\Prob{Y \geq (1 + z)\mu_n} \leq \exp(-z^2 \mu_n/3 + c_n)
	\end{equation*}
	the moment $\Expect{(1 + Y)^k}$, $k > 0$, satisfies for any $z > 0$ 
	\begin{equation*}
	\Expect{(1 + Y)^k} \leq \frac{2 n}{} 
	\end{equation*}
\end{lemma}

\section{Proofs}

Begin by expanding
\begin{align}
\label{eqn: expansion}
 \bigg| N \epsilon^d & \Expect{R_{XY}} - \Expect{\frac{2}{p(X) + q(X)} + \frac{2}{p(Y) + q(Y)}} \bigg| \nonumber \\
& = N \epsilon^d \abs{\Expect{R_{XY}} - \Expect{\frac{1}{d(X)} + \frac{1}{d(Y)}}} \nonumber \\
& +  N \epsilon^d \abs{\Expect{\frac{1}{d(X)} - \frac{1}{N \Prob{B(X,\epsilon)} } + \frac{1}{d(Y)} - \frac{1}{N \Prob{B(Y,\epsilon)}}}} \nonumber \\
& + \abs{ \Expect{\frac{\epsilon^d }{\Prob{B(X,\epsilon)} }  - \frac{2}{p(X) + q(X)} } + \Expect{ \frac{\epsilon^d }{\Prob{B(Y,\epsilon)} }  - \frac{2}{p(Y) + q(Y)} } } 
\end{align}

We will bound the summands on the right side of (\ref{eqn: expansion}) from last to first.

\paragraph{Third term.}

For the last term, we begin by rewriting
\begin{equation*}
\abs{ \frac{\epsilon^d }{\Prob{B(X,\epsilon)} }  - \frac{2}{p(X) + q(X)} } \leq \abs{ \frac{\epsilon^d(p(X) + q(X)) - 2 \Prob{B(X,\epsilon)}}{\Prob{B(X,\epsilon)}[p(X) + q(X)]} }
\end{equation*}

Then, we can bound the numerator using the fact we have required the densities $p$ and $q$ be Holder continuous, so
\begin{align*}
[p(X) + q(X)] \epsilon^d - 2 \Prob{B(X,\epsilon)} & = \int_{B(X,\epsilon)} [p(\x) - p(\z)] d\z + \int_{B(X,\epsilon)} [q(\x) - q(\z)] d\z \\
& \leq \int_{B(X,\epsilon)} 2 \norm{x - y}^{\alpha} d\z\\
& \leq 2 \epsilon^{\alpha + d}.
\end{align*}
We can lower bound the denominator using the lower bound on our densities
\begin{align*}
\Prob{B(X,\epsilon)}[p(X) + q(X)] \geq \epsilon^d(p_{\min} + q_{\min})^2 / 2
\end{align*}
and therefore
\begin{equation*}
\frac{\epsilon^d }{\Prob{B(X,\epsilon)} }  - \frac{2}{p(X) + q(X)} \leq \frac{4 \epsilon^{\alpha}}{(p_{\min} + q_{\min})^2}.
\end{equation*}

The same bound holds for the corresponding term with $Y$ instead of $X$.

\paragraph{Second term.}

To bound the second term, we will upper and lower bound $\Expect{ \frac{1}{d(X)} }$ by something close to $\Expect{ \frac{1}{N\Prob{B(X,\epsilon)}} }$.

The lower bound
\begin{align*}
\Expect{ \frac{1}{d(X)} } & = \Expect{ \Expect{\frac{1}{d(X)} \bigg| X} } \\
& \geq \Expect{ \frac{1}{1 + (N - 1)\Prob{B(X,\epsilon)}} } 
\end{align*}
follows from Jensen's inequality.

For the upper bound, note that the distribution of $d(X)$, conditional on $X$, is $1 + \text{Binomial}(N - 1,\Prob{B(X,\epsilon)})$. Then, letting $q = \Prob{B(X,\epsilon)}$
\begin{align*}
\Expect{\frac{1}{d(X)} \bigg| X} & = \sum_{k = 0}^{N - 1} \frac{1}{k + 1} {N - 1 \choose k} q^k (1 - q)^{N - 1 - k} \\
& = \frac{1}{Nq} \sum_{k = 0}^{N - 1}  {N - 1 \choose k + 1} q^{k + 1} (1 - q)^{N - 1 - k} \\
& \leq \frac{1}{Nq} \sum_{k = 0}^{N} {N \choose k } q^k (1 - q)^{N - k} \\
& = \frac{1}{Nq} \left(q + (1 - q)\right)^N = \frac{1}{Nq}.
\end{align*}

Combining this with the above, we have
\begin{align*}
N \epsilon^d \abs{\Expect{\frac{1}{d(X)} - \frac{1}{N \Prob{B(X,\epsilon)} } } } & \leq N \epsilon^d \abs{ \Expect{\frac{1}{1 + (N - 1)\Prob{B(X,\epsilon)}}}  - \Expect{\frac{1}{N\Prob(B(X, \epsilon))}}} \\
& \leq N \epsilon^d \abs{ \Expect{\frac{1}{N^2 \Prob{B(X, \epsilon)}^2}} }.
\end{align*}
with a corresponding bound holding for $Y$.

\paragraph{First term.}
We begin by reducing the first term to a product of moments and inverse moments of maxima and minima of binomials.
\begin{align*}
N \epsilon^d \abs{\Expect{R_{XY}} - \Expect{\frac{1}{d(X)} + \frac{1}{d(Y)}}} & \overset{(i)}{\leq} \frac{2 a_1}{\epsilon^2} \Expect{\frac{d^2_{\max}}{d^3_{\min}} \cdot (1 + 2 \frac{\widetilde{N}_{\max}}{\widetilde{N}_{\min}})} \\
& \overset{(ii)}{\leq} \frac{2 a_1}{\epsilon^2} \left(2 \Expect{d_{\max}^8} \cdot \Expect{ d_{\min}^{12} } \cdot \Expect{ \widetilde{N}_{\max}^8} \cdot \Expect{\frac{1}{\widetilde{N}_{\min}^8}} \right)^{1/4}\\
& + \frac{2 a_1}{\epsilon^2} \left( \Expect{d_{\max}^4} \cdot \Expect{d_{\min}^6} \right)^{1/4}
\end{align*}

where $(i)$ follows from Lemma \ref{lem: vonluxburg_poincare} and $(ii)$ from repeated applications of Holder's inequality.

\end{document}

### 2019-01-30 notes/graph_testing.tex
added backlog of notes

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\makeatletter
\newcommand{\leqnomode}{\tagsleft@true}
\newcommand{\reqnomode}{\tagsleft@false}
\makeatother

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Var}[1]{\mathrm{Var}\left( #1 \right)}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\dotp}[2]{\langle #1 , #2 \rangle}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\defeq}{\overset{\mathrm{def}}{=}}
\newcommand{\convweak}{\overset{w}{\rightharpoonup}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\Bin}{\mathrm{Bin}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\Bal}{\textrm{Bal}}
\newcommand{\Cut}{\textrm{Cut}}
\newcommand{\Ind}{\textrm{Ind}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Perp}{\perp \! \! \! \perp}
\newcommand{\Naturals}{\mathbb{N}}

\newcommand{\Linv}{L^{\dagger}}
\newcommand{\tr}{\text{tr}}
\newcommand{\h}{\textbf{h}}
% \newcommand{\l}{\ell}
\newcommand{\x}{\textbf{x}}
\newcommand{\y}{\textbf{y}}
\newcommand{\bl}{\bm{\ell}}
\newcommand{\bnu}{\bm{\nu}}
\newcommand{\Lx}{\mathcal{L}_X}
\newcommand{\Ly}{\mathcal{L}_Y}
\DeclareMathOperator*{\argmin}{argmin}

\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\A}{\mathcal{A}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\N}{\mathbb{N}}
\newcommand{\E}{\mathcal{E}}

%%% Matrix related notation
\newcommand{\Xbf}{\mathbf{X}}
\newcommand{\Ybf}{\mathbf{Y}}
\newcommand{\Zbf}{\mathbf{Z}}
\newcommand{\Abf}{\mathbf{A}}
\newcommand{\Dbf}{\mathbf{D}}
\newcommand{\Wbf}{\mathbf{W}}
\newcommand{\Lbf}{\mathbf{L}}
\newcommand{\Ibf}{\mathbf{I}}
\newcommand{\Bbf}{\mathbf{B}}

%%% Vector related notation
\newcommand{\lbf}{\bm{\ell}}
\newcommand{\fbf}{\mathbf{f}}

%%% Set related notation
\newcommand{\Dset}{\mathcal{D}}
\newcommand{\Aset}{\mathcal{A}}
\newcommand{\Wset}{\mathcal{W}}

%%% Distribution related notation
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Qbb}{\mathbb{Q}}
% \newcommand{\Pr}{\mathrm{Pr}}}

%%% Functionals
\newcommand{\1}{\mathbf{1}}
\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden} 
\newtheorem{definition}{Definition}[section]

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}

\theoremstyle{remark}
\newtheorem{remark}{Remark}

\begin{document}

\title{\textcolor{red}{Testing with Graphs}}
\author{Alden Green}
\date{\today}
\maketitle

\paragraph{Two-sample non-parametric hypothesis-testing problem.}

For fixed integers $n_1 + n_2 = n$, let $\Xbf = \set{x_1, \ldots, x_{n_1}} \subset \Rd$ and $\Ybf = \set{y_1, \ldots, y_{n_2}}$ be sampled i.i.d from distributions $\Pbb$ and $\Qbb$ with density functions $p$ and $q$, respectively, both with support on $D \subset \Rd$. Our statistical problem is testing the null hypothesis $H_0: \Pbb = \Qbb$ vs. the alternative $H_1: \Pbb \neq \Qbb$, where our knowledge of $\Pbb$ and $\Qbb$ come from the samples $\Xbf$ and $\Ybf$.

\paragraph{Integral Probability Metric.}
Given $\F$ a class of real-valued bounded measurable functions on $D$, the \emph{integral probability metric} between $\Pbb$ and $\Qbb$ with respect to $\F$ is
\begin{equation*}
\gamma_{\F}(\Pbb,\Qbb) = \sup_{f \in \F}\abs{\int_{D} f d\Pbb - \int_{D} f d\Qbb}
\end{equation*}

One such IPM that has not received close attention until surprisingly recently is the \emph{(weighted) Sobolev IPM}. For $f \in L^2(\Dset)$ and $\rho$ a density function over $\Dset$, the ($\rho^2$-weighted) Sobolev 1,2 norm of $f$ is given by
\begin{equation*}
\norm{f}_{1,2,\rho^2} := 
\begin{cases}
\int_{\Dset}\norm{\nabla_x f(x)}^2 \rho^2 dx,~ & \text{$f \in H^1(\Dset)$} \\
\infty,~ & \text{$f \in L^2(\Dset) \setminus H^1(\Dset)$}
\end{cases}
\end{equation*}
where $H^1(\Dset)$ is \textcolor{red}{the Sobolev space of $L^2(\Dset)$ functions with weak derivative $\nabla_x f(x) \in L^2(\Dset)$.}

Consider the unit ball of $\norm{\cdot}_{1,2,\rho^2}$,
\begin{equation*}
\Wset^{1,2}(\Dset,\rho^2) := \set{f: \norm{f}_{1,2,\rho^2} \leq 1}.
\end{equation*}
The weighted Sobolev IPM is simply $\gamma_{\Wset^{1,2}(\Dset,\rho^2)}(\Pbb,\Qbb)$.

\paragraph{Neighborhood graph.}

Let $G_{n,r_n} = (V,E,w)$ denote a weighted, undirected graph constructed from the samples $\Zbf = \set{z_1, \ldots, z_n} = (\Xbf,\Ybf)$ where $V = \set{1, \ldots, n}$, and $w_{uv} = K(z_u, z_v):= k(\frac{\norm{z_u - z_v}}{\epsilon_n}) \geq 0$ for $u,v \in V$, and a particular kernel function $k$. Here $(u,v) \in E$ if and only if $w_{uv} > 0$.

Motivated by the integral probability metric, our test statistics will be of the form
\begin{equation*}
\gamma_{\F_n}(\Pbb_n,\Qbb_n) = \sup_{f_n \in \F_n}\abs{\int_{D} f_n d\Pbb_n - \int_{D} f_n d\Qbb_n}
\end{equation*}
where
\begin{equation*}
\Pbb_n := \frac{1}{n_1}\sum_{i = 1}^{n_1} \delta_{x_i}, ~~ \Qbb_n := \frac{1}{n_2}\sum_{i = 1}^{n_2} \delta_{y_i}
\end{equation*}
are the empirical distributions of $\Xbf$ and $\Ybf$, respectively, and $\F_n$ is a \textcolor{red}{class of functions $f_n: \Zbf \to \Reals$ exhibiting some regularity with respect to the neighborhood graph $G_{n,r}$.}

\paragraph{Laplacian smoothing and Total Variation denoising.}

For convenience, we number the edges $E  = (e_1, \ldots, e_m)$. We denote by $\Bbf \in \Reals^{m \times n}$ the edge incidence matrix of $G_{n,r}$, which for $k$th edge $e_k = (u,v)$ has $k$th row $\Bbf_{k} = (0,\ldots,-w_{uv},\ldots,w_{uv},\ldots,0)$ with a $-w_{uv}$ in the $u$th location, and a $w_{uv}$ in the $v$th location. The random (unnormalized) Laplacian matrix is then $\Lbf = \Bbf^T \Bbf$. We also introduce a \emph{label vector}, given by $\lbf = (\ell_1, \ldots, \ell_n)$ with
\begin{equation}
\label{eqn: label_vector}
\ell_k = 
\begin{cases}
\frac{n}{n_1},~~~ \text{$z_k \in \Xbf$} \\
-\frac{n}{n_2}, ~ \text{$z_k \in \Ybf$}
\end{cases}
\end{equation}

Our test statistics $T_1(\lbf; G_{n,r})$ and $T_2(\lbf; G_{n,r})$ are defined as follows:
\begin{align*}
	T_1(\lbf; G_{n,r}) := \sup_{\fbf \in \Reals^n: \norm{\Bbf \fbf}_1 \leq C_{n,r}} ~ \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_k \\
	T_2(\lbf; G_{n,r}) := \sup_{\fbf \in \Reals^n: \norm{\Bbf \fbf}_2^2 \leq C_{n,r}} ~ \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_k
\end{align*}
where \textcolor{red}{$C_{n,r} = $} and we write $\fbf = (f_1,\ldots, f_n)$. We note that, as promised, these satisfy the form
\begin{align*}
T_1(\lbf; G_{n,r}) & = \sup_{f_n \in TV_n}\abs{\int_{D} f_n d\Pbb_n - \int_{D} f_n d\Qbb_n} \\
T_2(\lbf; G_{n,r}) & = \sup_{f_n \in \Wset_n}\abs{\int_{D} f_n d\Pbb_n - \int_{D} f_n d\Qbb_n}
\end{align*}
where $TV_n = \set{\fbf: \norm{\Bbf \fbf}_1 \leq C_{n,r}}$ and $\Wset_n = \set{\fbf: \norm{\Bbf \fbf}_2 \leq C_{n,r}}$.

\section{Consistency under fixed alternative}

\paragraph{Binomialized data model.}
For technical reasons, we would like $z_1, \ldots, z_n$ to be independent and identically distributed. We consider the following generative model, which we term the \emph{binomialized data model}:

Fix $n \in \Naturals > 0$, and $n_1 \sim \Bin(n,1/2),~ n_2 = n - n_1$. Then, let $x_1, \ldots, x_{n_1} \in \Rd$ be a sequence of i.i.d random points chosen according to $\Pbb$, and $y_1, \ldots, y_{n_2} \in \Rd$ a separate sequence of i.i.d random points chosen according to $\Qbb$, with $x_j \perp y_k$ for all $j,k$. Fix $\widetilde{\Zbf} = (\widetilde{z}_1,\ldots,\widetilde{z}_n) := (x_1, \ldots, x_{n_1}, y_1, \ldots, y_{n_2})$.  Finally, for a permutation $\pi: [n] \to [n]$ chosen uniformly at random among all such permutations, let $\Zbf = (z_1,\ldots,z_n) =  (\widetilde{z}_{\pi(1)}, \ldots, \widetilde{z}_{\pi(n)})$.

The label vector $\lbf$ remains defined as in \eqref{eqn: label_vector} with respect to $\Zbf$. Note that now $z_i \overset{i.i.d}{\sim} \frac{\Pbb}{2} + \frac{\Qbb}{2}$, as we desired, with density function $\mu(x) := \frac{p(x) + q(x)}{2}$.

\begin{theorem}
	\label{thm: LLN_for_t2}
	Let $d \geq 2$ and let $\Dset \subset \Rd$ be an open, bounded, connected set with Lipschitz boundary. Let $\mu$ satisfy
	\begin{equation*}
	m \leq \mu(x) \leq M \tag{$\forall x \in D$}
	\end{equation*}
	for some $0 < m \leq M$. Let $(r_n)$ be a sequence of positive numbers converging to $0$ and satisfying
	\begin{align*}
	& \lim_{n \to \infty} \frac{(\log n)^{3/4}}{n^{1/2}} \frac{1}{r_n} = 0 ~~ \text{if $d = 2$} \\
	& \lim_{n \to \infty} \frac{(\log n)^{1/d}}{n^{1/d}} \frac{1}{r_n} = 0 ~~ \text{if $d \geq 3$}
	\end{align*}
	Assume the kernel $k$ satisfies conditions: 
	\begin{align*}
	& k(0) > 0 ~ \text{and $k$ is continuous at $0$.} \tag{\textbf{K1}} \\
	& k~ \text{is non-increasing.} \tag{\textbf{K2}} \\
	& \text{The integral $\int_{0}^{\infty} k(r) r^{d+1} dr$ is finite.} \tag{\textbf{K3}}
	\end{align*}

	Then with probability one the following statement holds: For $(z_1, \ldots, z_n)$ chosen under the binomialized data model
	\begin{equation*}
	\lim_{n \to \infty} T_2(\lbf; G_{n,r_n}) = \gamma_{\Wset^{1,2}(\Dset,\mu^2)}(\Pbb,\Qbb).
	\end{equation*}
\end{theorem}

\section{Proofs}

Fix $\mu_n = \frac{d\Pbb_n + d\Qbb_n}{2}$. \textcolor{red}{We will show a variational form of convergence of $\mu_n$ to $\mu$.}

\subsection{Gamma convergence of constraint}

\begin{definition}[$TL^2$ convergence]
	Denote by $\mathfrak{B}(\Dset)$ the Borel $\sigma$-algebra of $\Dset$ and $\mathcal{P}(\Dset)$ the set of all Borel probability measures on $\Dset$. Given a Borel map $T: \Dset \to \Dset$, the \emph{push-forward} of $\mu$ by $T$ is given by
	\begin{equation*}
	T_{\star}\mu(\Aset) := \mu(T^{-1}(\Aset)), ~~ \Aset \in \mathfrak{B}(\Dset).
	\end{equation*}

	Given $\widetilde{\mu} \in \mathcal{P}(\Dset)$, we say that $T$ is a \emph{transportation map} between $\mu$ and $\widetilde{\mu}$ if $T_{\star}\mu = \widetilde{\mu}$. If for a sequence $(T_n)$ of transportation maps
	\begin{equation*}
	\int_{\Dset} \abs{x - T_n(x)}^2 d\mu(x) \to 0,~ \text{as $n \to \infty$}
	\end{equation*}
	we refer to the sequence as \emph{stagnating.}

	Take $f \in L^2(\mu)$ and a sequence $(f_n)$ with $f_n \in L^2(\mu_n)$ for $n = 1,2,\ldots$. If there exists a stagnating sequence of transportation maps $(T_n)$ such that
	\begin{equation*}
	\int_{\Dset} \abs{ f - T_n(f_n(x)) }^2 d\mu(x) \to 0,~ \text{as $n \to \infty$}
	\end{equation*}
	we say that $(f_n)$ converges $TL^2$ to $f$, and write $f_n \overset{TL^2}{\to} f$.
\end{definition}

We restate Theorem 1.4 of \cite{trillos2018}, changing notation to match the rest of this paper.

\begin{theorem}[Theorem 1.4 of \cite{trillos2018}]
	Under the setup and conditions of Theorem \ref{thm: LLN_for_t2}, with probability one the following statements hold:
	\begin{itemize}
		\item \emph{Liminf inequality:}
		For all $f \in L^2(\mu)$ and all sequences $(f_n)$ with $f_n \in L^2(\mu_n)$ and $f_n \overset{TL^2}{\to} f$,
		\begin{equation*}
		\liminf_{n \to \infty} \frac{1}{n^2 r_n^{d + 2}} \norm{Bf}_2^2 \geq \textcolor{red}{\sigma_{k}} \norm{f}_{1,2,\rho^2}
		\end{equation*}

		\item \emph{Limsup inequality:} For all $f \in L^2(\mu)$, there exists a sequence $(f_n)$ with $f_n \in L^2(\mu_n)$ and $f_n \overset{TL^2}{\to} f$ such that
		\begin{equation*}
		\limsup_{n \to \infty} \frac{1}{n^2 r_n^{d + 2}} \norm{Bf}_2^2 \leq \textcolor{red}{\sigma_{k}} \norm{f}_{1,2,\rho^2}
		\end{equation*}

		\item \emph{Compactness property:} Every sequence $(f_n)$ with $f_n \in L^2(\mu_n)$ satisfying
		\begin{equation*}
		\sup_{n \in \Naturals} \frac{1}{n^2 r_n^{d + 2}} \norm{Bf}_2^2 < \infty
		\end{equation*}
		is precompact in $TL^2$, that is, every subsequence of $(f_n)$ has a further subsequence which converges in the $TL^2$-sense to an element of $L^2(\Dset)$.
	\end{itemize}

\end{theorem}

\subsection{Continuity of risk functional}

\subsection{Proof of Theorem \ref{thm: LLN_for_t2}}

\clearpage

\bibliography{../graph_testing_bibliography}
\bibliographystyle{plain}

\end{document}

### 2019-01-30 notes/weekly_notes/12.31.18.NotesForWeek.tex
added backlog of notes

@@
 \newcommand{\Perp}{\perp \! \! \! \perp}
+\newcommand{\Naturals}{\mathbb{N}}
 
@@
 
+%%% Set related notation
+\newcommand{\Dset}{\mathcal{D}}
+
+%%% Distribution related notation
+\newcommand{\Pbb}{\mathbb{P}}
+\newcommand{\Qbb}{\mathbb{Q}}
+% \newcommand{\Pr}{\mathrm{Pr}}}
+
+%%% Functionals
+\newcommand{\1}{\mathbf{1}}
 \newtheoremstyle{alden}
@@
 
-We are given two unknown distributions, $P$ and $Q$, both supported on $D \subset \Rd$, with continuous density functions $p$ and $q$, respectively.
+We are given two unknown distributions, $\Pbb$ and $\Qbb$, both supported on \textcolor{red}{$D \subset \Rd$}, with \textcolor{red}{continuous} density functions $p$ and $q$, respectively.
 We have the capacity to sample from either distribution.
-Our goal is to test the nonparametric hypothesis $H_0: P = Q$ vs.
-the alternative $H_1: P \neq Q$.
+Our goal is to test the nonparametric hypothesis $H_0: \Pbb = \Qbb$ vs.
+the alternative $H_1: \Pbb \neq \Qbb$.
 
-Under the \textbf{binomial data model}, we sample data $\set{z_1, \ldots, z_n}$ as follows.
-For each $i$, we draw an independent Rademacher label $\ell_i \in \set{1, -1}$, $\mathbb{P}(\ell_i = 1) = \mathbb{P}(\ell_i = -1) = 1/2$.
-Then, if $\ell_i = 1$ we sample $z_i \sim P$, whereas if $\ell_i = -1$ we samle $z_i \sim Q$ otherwise.
-Define $1_X$ to be the length $n$ indicator vector for $\ell_i = 1$
+Under the \emph{binomial data model}, we sample data $\set{z_1, \ldots, z_n}$ as follows: for $i = 1,\ldots,n$, we draw an independent Rademacher label $\ell_i \in \set{1, -1}$, $\Pr(\ell_i = 1) = \Pr(\ell_i = -1) = 1/2$.
+Then, if $\ell_i = 1$ we sample $z_i \sim \Pbb$, whereas if $\ell_i = -1$ we sample $z_i \sim \Qbb$.
+Define $\1_X$ to be the length $n$ indicator vector for $\ell_i = 1$
 \begin{equation*}
-1_X(i) = 
+\1_X(i) = 
 \begin{cases}
@@
 \end{equation*}
-and similarly for $1_Y$
+and similarly for $\1_Y$
 \begin{equation*}
-1_Y(i) = 
+\1_Y(i) = 
 \begin{cases}
@@
 \end{equation*}
-Denote the number of positive labels $N = \sum_{i = 1}^{n} 1_x(i)$, and let $M$ be such that $n = N + M$.
-Then, let the \textbf{normalized label vector} $a$ be given by $a = \frac{1}{N}1_X - \frac{1}{M} 1_Y$.
+Denote the number of positive labels $N = \sum_{i = 1}^{n} \1_X(i)$, for $M$ the number of negative labels we have $n = N + M$.
+Then the \emph{normalized label vector} $a$ is given by $a = \frac{1}{N}\1_X - \frac{1}{M}\1_Y$.
 
@@
 
-Heuristically, our graph based test will attempt to identify whether samples from the same distribution are on average more similar than samples from different distributions.
+Heuristically, our graph-based test will attempt to identify whether samples from the same distribution are on average more similar than samples from different distributions.
 As a result, we introduce $K$, a \textbf{kernel function} which measures similarity, and make the following assumptions on $K$
@@
 \begin{equation*}
-\mathcal{E}_n^2(u_n) \defeq \frac{1}{n^2 \epsilon_n^{d+1}} \sum_{i = 1}^{n} \sum_{j = 1}^{n} K\left(\frac{\norm{z_i - z_j}}{\epsilon_n}\right)(u_n(z_i) - u_n(z_j))^2.
+\mathcal{E}_n^2(u_n) \defeq \frac{1}{n^2 \epsilon_n^{d+2}} \sum_{i = 1}^{n} \sum_{j = 1}^{n} K\left(\frac{\norm{z_i - z_j}}{\epsilon_n}\right)(u_n(z_i) - u_n(z_j))^2.
 \end{equation*}
@@
 
-\begin{proof}[Proof attempt of Theorem \ref{thm: consistency_of_empirical_risk_minimizer}]
+\begin{proof}[Proof of Theorem \ref{thm: consistency_of_empirical_risk_minimizer}]
+	We know $\E_{\infty}^2(u^{\star}) \leq 1$ (otherwise $R^{(con)}(0) = 0 \leq R^{(con)}(u^{\star})$)
 
-We begin by restricting ourselves to the case where $\underset{n \to \infty}{\limsup}~ \E_n^2(u_n^{\star}) \leq 1$.
-As a result of Lemma 
+	By Lemma \ref{thm: garciatrillos17}, we have that there exists some $u_n \overset{TL^2}{\to} u^{\star}$ such that
+	\begin{equation*}
+	\limsup_{n \to \infty} \E_n^2(u_n) \leq \E_{\infty}^2(u^{\star}) \leq 1
+	\end{equation*}
+	and therefore by Lemma \ref{lem: conv_of_risk_functional}
+	\begin{equation*}
+	\limsup_{n \to \infty} R_n(u_n) = R(u^{\star})
+	\end{equation*}
+	Of course, we do not know that $\E_n^2(u_n) \leq 1$, and so we do not know that the $R_n^{(con)}(u_n) < \infty$, even in the limit.
+However, taking $u_n' = u_n \cdot \left(\max\{1, \E_n^2(u_n) \}\right)^{-1}$, we have $\E_n^2(u_n') \leq 1$.
+Moreover,
+	\begin{align}
+	\label{eqn: theorem_consistency_1}
+	\lim_{n \to \infty} R_n(u_n') & = \lim_{n \to \infty} \frac{R_n(u_n)}{\bigl(\max\left\{1, \E_n^2(u_n)\right\}\bigr)} \nonumber \\
+	& = R(u^{\star})
+	\end{align}
+	where the latter equality follows from the continuous mapping theorem. 
+	Since
+	\begin{equation*}
+	 R_n(u_n^{\star}) \leq R_n^{(con)}(u_n^{\star}) \leq R_n^{(con)}(u_n') = R_n(u_n')
+	\end{equation*}
+	we have
+	\begin{equation*}
+	\lim_{n \to \infty} R_n(u_n^{\star}) \leq R(u^{\star}).
+	\end{equation*}
+
+	Finally, the above reasoning implies
+	\begin{equation*}
+	\lim_{n \to \infty} R_n^{(con)}(u_n^{\star}) \leq R(u^{\star}) < \infty. 
+	\end{equation*}
+	As a result, clearly
+	\begin{equation*}
+	\limsup_{n \to \infty} \E_n^2(u_n^{\star}) \leq 1 < \infty 
+	\end{equation*}
+	and so by Theorem \ref{thm: garciatrillos17}, we have that every subsequence of $u_n^{\star}$ is $TL^2$ convergent. As a result, 
+	\begin{equation*}
+	\liminf_{n \to \infty} \E_n^2(u_n^{\star}) \geq \inf_{u \in L^2(\nu)} R(u) = R(u^{\star})
+	\end{equation*}
+	and so we have shown
+	\begin{equation*}
+	\lim_{n \to \infty} R_n(u_n^{\star}) = R(u^{\star}).
+	\end{equation*}
 \end{proof}
 \subsection{Technical results.}
+
+\begin{theorem}[\textcolor{red}{Garcia-Trillos 17}]
+	\label{thm: garciatrillos17}
+	Let $d \geq 2$ and let $\Dset \subset \Rd$ be an open, bounded, connected set with Lipschitz boundary. Let $\mu$ be a probability measure on $\Dset$ with continuous density $\rho$, satisfying 
+	\begin{equation*}
+	m \leq \rho(x) \leq M \tag{$\forall x \in D$}
+	\end{equation*}
+	for some $0 < m \leq M$. Let $z_1, \ldots, z_n$ be a sequence of i.i.d random points chosen according to $\mu$. Let $(\epsilon_n)$ be a sequence of positive numbers converging to $0$ and satisfying
+	\begin{align*}
+	& \lim_{n \to \infty} \frac{(\log n)^{3/4}}{n^{1/2}} \frac{1}{\epsilon_n} = 0 ~~ \text{if $d = 2$} \\
+	& \lim_{n \to \infty} \frac{(\log n)^{1/d}}{n^{1/d}} \frac{1}{\epsilon_n} = 0 ~~ \text{if $d \geq 3$}
+	\end{align*}
+	Assume the kernel $K$ satisfies conditions: 
+	\begin{align*}
+	& K(0) > 0 \text{and $K$ is continuous at $0$.} \tag{\textbf{K1}} \\
+	& K \text{is non-increasing.} \tag{\textbf{K2}} \\
+	& \text{The integral $\int_{0}^{\infty} K(r) r^{d+1} dr$ is finite.} \tag{\textbf{K3}}
+	\end{align*}
+	Then, with probability one, the following statement holds:
+	\begin{equation*}
+	\E_n^2(u_n) \overset{\Gamma}{\to} \E_{\infty}^2(u)
+	\end{equation*}
+	in the $TL^2$ sense.
+
+	Moreover, every sequence $(u_n)$ with $u_n \in L^2(\mu_n)$ for which
+	\begin{align*}
+	\sup_{n \in \Naturals} \norm{u_n}_{\mu_n} & < \infty \\
+	\sup_{n \in \Naturals} \E_n^2(u_n) & < \infty
+	\end{align*}
+	is pre-compact in $TL^2$.
+\end{theorem}
 

### 2019-02-08 notes/graph_testing.tex
added notes for this week

@@
 \newcommand{\1}{\mathbf{1}}
+
 \newtheoremstyle{alden}
@@
 \end{align*}
-where \textcolor{red}{$C_{n,r} = $} and we write $\fbf = (f_1,\ldots, f_n)$. We note that, as promised, these satisfy the form
+where $C_{n,r} = \frac{\sigma_k}{n^2 r_n^{d+2}}$ and we write $\fbf = (f_1,\ldots, f_n)$. We note that, as promised, these satisfy the form
 \begin{align*}
@@
 
-\begin{theorem}
+\begin{theorem}[Pointwise limit of Laplacian smooth test statistic.]
 	\label{thm: LLN_for_t2}
@@
 
-	Then with probability one the following statement holds: For $(z_1, \ldots, z_n)$ chosen under the binomialized data model
+	Then with probability one the following statement holds: For $(z_1, \ldots, z_n)$ chosen under the binomialized data model,
 	\begin{equation*}
@@
 \begin{theorem}[Theorem 1.4 of \cite{trillos2018}]
+	\label{thm: gamma_convergence_energy_functional}
 	Under the setup and conditions of Theorem \ref{thm: LLN_for_t2}, with probability one the following statements hold:
@@
 		\begin{equation*}
-		\limsup_{n \to \infty} \frac{1}{n^2 r_n^{d + 2}} \norm{Bf}_2^2 \leq \textcolor{red}{\sigma_{k}} \norm{f}_{1,2,\rho^2}
+		\limsup_{n \to \infty} \frac{1}{n^2 r_n^{d + 2}} \norm{\Bbf \fbf_n }_2^2 \leq \textcolor{red}{\sigma_{k}} \norm{f}_{1,2,\rho^2}
 		\end{equation*}
+		where $\fbf_n = (f_n(z_1), \ldots, f_n(z_n))$.
 
@@
 
+\begin{lemma}
+	\label{lem: continuity_of_risk_functional}
+	With probability one the following statement holds: If a sequence $(f_n)$ where $f_n: \set{z_1, \ldots,z_n} \to [-1,1]^{n}$ converges $TL^2$ to $f \in L^2(\mu)$, then
+	\begin{equation*}
+	\lim_{n \to \infty} \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_n(z_k) = \int_{\Dset} f d\Pbb(x) - \int_{\Dset} fd\Qbb(x).
+ 	\end{equation*}
+\end{lemma}
+
 \subsection{Proof of Theorem \ref{thm: LLN_for_t2}}
+
+Let $f^{\star}$ be the witness function for $\gamma_{\Wset^{1,2}(\Dset,\mu^2)}(\Pbb,\Qbb)$, meaning
+\begin{equation*}
+\abs{\int_{\Dset}f^{\star}d\Pbb - \int_{\Dset}f^{\star}d\Qbb} = \gamma_{\Wset^{1,2}(\Dset,\mu^2)}(\Pbb,\Qbb)
+\end{equation*}
+where $f^{\star} \in \Wset^{1,2}(\Dset,\mu^2)$ implies $f^{\star} \in L^2(\mu)$.
+By the limsup inequality in Theorem \ref{thm: gamma_convergence_energy_functional}, there exists some $(f_n) \overset{TL^2}{\to} f$ with $f_n \in L^2(\mu_n)$ such that
+\begin{equation}
+\label{eqn: limsup_discrete_energy}
+\limsup_{n \to \infty} \frac{1}{n^2 r_n^{d+2}} \norm{\Bbf \fbf_n}_2^2 \leq \textcolor{red}{\sigma_k} \norm{f}_{1,2,\mu^2} \leq \sigma_k.
+\end{equation} 
+From Lemma \ref{lem: continuity_of_risk_functional}
+\begin{equation*}
+\lim_{n \to \infty} \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_n(z_k) = \int_{\Dset} f d\Pbb(x) - \int_{\Dset} f d\Qbb(x) = \gamma_{\Wset^{1,2}(\Dset,\mu^2)}.
+\end{equation*}
+and along with \eqref{eqn: limsup_discrete_energy}, this implies
+\begin{equation*}
+\lim_{n \to \infty} T_2(\lbf; G_{n,r}) \leq \gamma_{\Wset^{1,2}(\Dset,\mu^2)}(\Pbb,\Qbb).
+\end{equation*}
+
+Let $\Lbf^{\dagger}$ be the pseudoinverse of the Laplacian matrix $\Lbf$. We introduce $f_n^{\star} = \lbf^T \Lbf^{\dagger} \lbf$, which satisfies
+\begin{equation*}
+\abs{\int_{\Dset}f_n^{\star}d\Pbb_n - \int_{\Dset}f_n^{\star}d\Qbb_n} = T_2(\lbf; G_{n,r})
+\end{equation*}
+Note that
+\begin{equation*}
+\norm{\Bbf \fbf_n^{\star}}_2^2 \leq C_n \Longrightarrow \frac{1}{n^2 r_n^{d+2}} \norm{\Bbf \fbf_n^{\star}}_2^2  \leq \sigma_k < \infty
+\end{equation*}
+and so by the compactness property in Theorem \ref{thm: gamma_convergence_energy_functional}, every subsequence of $f_n^{\star}$ has a further subsequence which is $TL^2$-convergent.
+For simplicity, and without loss of generality, let us work along a subsequence which is convergent (and call it $f_n^{\star}$) to $f \in L^2(\mu)$.
+Then
+\begin{equation}
+\label{eqn: lower_bound_T2_1}
+\lim_{n \to \infty} \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_n^{\star}(z_k) = \abs{\int_{\Dset}f d\Pbb(x) - \int_{\Dset}f d\Qbb(x)}
+\end{equation}
+and by the liminf inequality of Theorem \ref{thm: gamma_convergence_energy_functional},
+\begin{equation*}
+\liminf_{n \to \infty} \frac{1}{n^2 r_n^{d+2}} \norm{\Bbf \fbf_n}_2^2 \geq \sigma_k \norm{f}_{1,2,\mu^2}
+\end{equation*}
+which implies $\norm{f}_{1,2,\mu^2} \leq 1$. This, along with \eqref{eqn: lower_bound_T2_1}, implies
+\begin{equation*}
+\lim_{n \to \infty} \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_n^{\star}(z_k) \geq \gamma_{\Wset^{1,2}(\Dset,\mu^2)}(\Pbb,\Qbb)
+\end{equation*}
 

### 2019-02-08 notes/reference_notes/padilla18.tex
added notes for this week

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Var}[1]{\mathrm{Var}\left( #1 \right)}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Matrices
\newcommand{\Xbf}{\mathbf{X}}

%%% Sets
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}

\title{Notes on 'Adaptive Non-Parametric Regression With the \emph{K}-NN Fused Lasso'}
\author{Alden Green}
\date{\today}
\maketitle

Let $\Xbf = x_1, \ldots, x_n$ be sampled i.i.d from $\mu$ with density function $p(\cdot)$ over some subset $\Xset$ of Euclidean space, and suppose
\begin{equation*}
y_i = f_0(x_i) + \epsilon_i, ~~~\epsilon_i \overset{i.i.d}{\sim}SG(\sigma^2)
\end{equation*}
holds for some unknown $f_0$. Let $\widehat{\theta}$ be the solution to the \emph{fused lasso}
\begin{equation*}
\widehat{\theta} := \argmin_{\theta \in \Reals^n} \set{\frac{1}{2} \norm{y - \theta}_2^2 + \lambda \norm{\nabla_G \theta}_1} 
\end{equation*}
where $\lambda > 0$ is a tuning parameter, and $\nabla_G$ is an oriented incidence matrix of the graph $G$.

The \emph{$K$-NN-FL} estimator computes the fused lasso over the $K$-NN graph $G_K$ of $\Xbf$. The \emph{$\epsilon$-FL} estimator computes the fused lasso over the $\epsilon$ graph $G_{\epsilon}$.

The assumptions required for Theorems \ref{thm: padilla18_theorem_1} and \ref{thm: padilla18_theorem_2} are as follows.
\begin{enumerate}[(a)]
	\item For all $x \in \Xset$
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty
	\end{equation*}
	\item The base measure $\mu$ in $\Xset$ satisfies
	\begin{equation*}
	r^d c_{1,d} \leq \mu(B_r(x)) \leq c_{2,d} r^d \tag{$\forall x \in \Xset$}
	\end{equation*}
	\item There exists a homeomorphism (continuous bijection with continuous inverse) $h: \Xset \to [0,1]^d$ such that
	\begin{equation*}
	L_{\min} d_{\Xset}(x,x') \leq \norm{h(x) - h(x')}_2  \leq L_{\max} d_{\Xset}(x,x') \tag{$\forall x,x' \in \Xset$}
	\end{equation*}
	\item The function $g_0 := f_0 \circ h^{-1}$ has bounded variation, meaning
	\begin{equation*}
	\abs{g_0}_{BV(\Omega)} := \sup \set{\int_{\Omega} g_0(x) \dive(g)(x)dx; g \in C_c^1(\Omega, \Rd), \norm{g}_{\infty} \leq 1} < \infty
	\end{equation*}
	where $\mathfrak{B}$ is the Borel $\sigma$-algebra of $\Omega = [0,1]^d$, $C_c^1(\Omega, \Rd)$ is the set of $\mathfrak{B}$-measurable, continuously differentiable functions into $\Rd$, and
	\begin{equation*}
	\norm{g}_{\infty} := \norm{\left(\sum_{i=1}^{d}g_i^2\right)^{1/2}}_{L_{\infty}(\Omega)}.
	\end{equation*}
	\item $g_0$ is piecewise Lipschitz\footnote{Technically, the requirement is slightly weaker than piecewise Lipschitz.}, meaning there exists a set $\Sset \subset (0,1)^d$ such that
	\begin{enumerate}
		\item $\nu(\Sset) = 0$.
		\item $\mu\Bigl( h^{-1}\bigl(S_{\epsilon} \cup ([0,1]^d \setminus \Omega_{\epsilon})\bigr) \Bigr) \leq C_{\Sset} \epsilon$
		\item There exists a positive constant $L_0$ such that if $z$ and $z'$ belong to the same connected component of $\Omega_{\epsilon} \setminus B_{\epsilon}(\Sset)$, then 
		\begin{equation*}
		\abs{g(z) - g(z')} \leq L_0 \norm{z - z'}_2
		\end{equation*}
	\end{enumerate}
	where $\Omega_{\epsilon} = [0,1]^d \setminus B_{\epsilon}(\partial [0,1]^d)$.
\end{enumerate}

\begin{theorem}
	\label{thm: padilla18_theorem_1}
	Let $K \asymp \log^{1 + 2r} n$ for some $r > 0$, Then under Assumptions \textcolor{red}{1-3}, with an appropriate choice of the tuning parameter $\lambda$, the \textcolor{red}{$K$-NN-FL} estimator $\widehat{\theta}$ satisfies
	\begin{equation*}
	\norm{\widehat{\theta} - \textcolor{red}{\theta^{\star}}}_n^2 = O_{\Pbb}\left(\frac{\log^{1 + 2r} n}{n} + \frac{\log^{1.5 + r} n}{n}\norm{\nabla_{G_K} \theta^{\star}}_1 \right)
	\end{equation*}

	This upper bound also holds for $\epsilon$-NN-FL if we replace $\norm{\nabla_{G_K} \theta^{\star}}_1 $ with $\norm{\nabla_{G_{\epsilon}} \theta^{\star}}_1$.
\end{theorem}

\begin{theorem}
	\label{thm: padilla18_theorem_2}
	Under Assumptions \textcolor{red}{1-5}, with an appropriate choice of the tuning parameter $\lambda$, the \textcolor{red}{$K$-NN-FL} estimator $\widehat{\theta}$ satisfies
	\begin{equation*}
	\norm{\widehat{\theta} - \textcolor{red}{\theta^{\star}}}_n^2 = \widetilde{O}_{\Pbb}\left( \frac{1}{n^{1/d}} \right).
	\end{equation*}
\end{theorem}

\section{Proof of Theorem \ref{thm: padilla18_theorem_1}}

To ease proofs, we will assume $\Xset = [0,1]^d$.

Construct $G_{lat} = (V_{lat}, E_{lat})$ a lattice graph with equal side lengths in $[0,1]^d$, where
\begin{align*}
V_{lat} & = P_{lat}(N) := \set{\left(\frac{i_1}{N} - \frac{1}{2N}, \cdots, \frac{i_d}{N} - \frac{1}{2N}\right): i_1, \ldots, i_d \in \set{1, \ldots, N}} \\
(z,z') & \in E_{lat}~ \text{if and only if $\norm{z - z'} \leq \frac{1}{N}$}
\end{align*}
where $z$ and $z' \in P_{lat}(N)$.

Denoting $I = P_{lat}$, we define
\begin{equation*}
P_I(x) = \argmin \set{\norm{x - z'}_{\infty}, z' \in P_{lat}(N)}
\end{equation*}

Then, let $C(z) = \set{x \in [0,1]^d: z =  P_I(x)}$ be the collection of cells associated with the mesh $P_{lat}(N)$, noting that $\set{C(z): z \in P_{lat}(N)}$ defines a partition over $[0,1]^d$.

\paragraph{Quantization.}

For a given $\theta \in \Reals^n$, the \emph{quantization} $\theta_I \in \Reals^n$
\begin{equation*}
(\theta_{I})_i := \theta_j,~ \text{ where }~ x_j = \argmin_{x_l, l \in [n]} \norm{P_I(x_i) - x_l}_{\infty}
\end{equation*}
is constant over every cell $C(z)$. We now induce a signal in $\Reals^{N^d}$ corresponding to the elements in $I$. Let $\set{z_1, \ldots, z_{N^d}} = I$. Then
we write
\begin{equation*}
I_k = \set{i \in [n]: P_I(x_i) = z_k} 
\end{equation*}
for $k = 1,\ldots, N^d$. Define $\theta^{I} \in \Reals^{N^d}$ by
\begin{equation*}
(\theta^{I})_k := 
\begin{cases}
(\theta_I)_i, x_i \in I_k \\
0, I_k = \emptyset
\end{cases}
\end{equation*}
where we note that $(\theta^{I})$ is well-defined since $(\theta_I)_i = (\theta_I)_j$ if $x_i$ and $x_j$ are both in $I_k$.

\subsection{Controlling counts of mesh}

Define the event $\Omega$ as: ``If $x_i \in C(z_k)$ and $x_i \in C(z_l)$ for $z_k, z_l \in I$ with $\norm{z_k - z_l}_2 \leq \frac{1}{N}$, then $x_i$ and $x_j$ are connected in the $K$-NN graph.'' Then,
\begin{lemma}
	\label{lem: mesh_counts}
	Take Assumptions \textcolor{red}{1-3}, and additionally assume that $N$ in the construction of $G_{lat}(N)$ is chosen as
	\begin{equation}
	\label{eqn: N_lower_bound}
	N \geq \left[\frac{3\sqrt{d}(2 c_{2,d} p_{\max})^{1/d} n^{1/d}}{L_{\min} K^{1/d}}\right].
	\end{equation}
	Then,
	\begin{equation*}
	\Pbb(\Omega) \geq 1 - n \exp(-K / 3).
	\end{equation*}
\end{lemma}

\subsection{Bounding Empirical Process}

\begin{lemma}

\end{lemma}

\subsection{Mesh embedding for $K$-NN graph}

\begin{lemma}
	Fix $N$ to satisfy \eqref{eqn: N_lower_bound}, and let us assume that the event $\Omega$ from Lemma \ref{lem: mesh_counts} holds. Denote $I = P_{lat}(N)$ to be the mesh. Then, for all $e \in \Reals^n$, it holds that
	\begin{equation*}
	\abs{e^T(\theta - \theta_I)} \leq 2 \norm{e}_{\infty} \norm{\nabla_{G_K} \theta}_1, \tag{$\forall \theta \in \Reals^n$}
	\end{equation*}
	Moreover, 
	\begin{equation*}
	\norm{D \theta^I}_1 \leq \norm{\nabla_{G_K} \theta}_1, \tag{$\forall \theta \in \Reals^n$}
	\end{equation*}
	where $D$ is the incidence matrix of $G_{lat}$.
\end{lemma}

\end{document}

### 2019-02-08 notes/weekly_notes/2.6.19.NotesForWeek.tex
added notes for this week

--- 2019-01-30 notes/graph_testing.tex
@@
 \newcommand{\Qbb}{\mathbb{Q}}
+\newcommand{\Gbb}{\mathbb{G}}
 % \newcommand{\Pr}{\mathrm{Pr}}}
@@
 
-\title{\textcolor{red}{Testing with Graphs}}
+\title{Notes for Week 2/1/18 - 2/8/18}
 \author{Alden Green}
@@
 \maketitle
-
-\paragraph{Two-sample non-parametric hypothesis-testing problem.}
 
@@
 
-\paragraph{Integral Probability Metric.}
-Given $\F$ a class of real-valued bounded measurable functions on $D$, the \emph{integral probability metric} between $\Pbb$ and $\Qbb$ with respect to $\F$ is
+Recall the \emph{Laplacian smooth} and \emph{total variation smooth} test statistics
+\begin{align*}
+T_1(\lbf; G_{n,r}) & = \sup_{\theta: \norm{\Bbf \theta}_1 \leq C_{n,r}}\abs{\lbf^T \theta} \\
+T_2(\lbf; G_{n,r}) & = \sup_{\theta: \norm{\Bbf \theta}_2 \leq C_{n,r}}\abs{\lbf^T \theta}
+\end{align*}
+where $\Bbf$ is the incidence matrix of the $r$-neighborhood graph and $C_{n,r} = \frac{\sigma_k}{n^2 r_n^{d+2}}$.
+
+Theorem \ref{thm: weak_convergence_null_hypothesis_T2} is the type of theorem we are looking for.
+
+\begin{theorem}
+	\label{thm: weak_convergence_null_hypothesis_T2}
+	Under \textcolor{red}{assumptions}, 
 	\begin{equation*}
-\gamma_{\F}(\Pbb,\Qbb) = \sup_{f \in \F}\abs{\int_{D} f d\Pbb - \int_{D} f d\Qbb}
+	\sqrt{n} T_2(\lbf; G_{n,r}) \rightsquigarrow \textcolor{red}{???}
+	\end{equation*}
+	under the null hypothesis $H_0: \Pbb = \Qbb$.
+\end{theorem}
+
+The rest of this document details the strategy for proving this convergence.
+
+\section{Quantization}
+
+To ease proofs, we will assume $\Dset = [0,1]^d$.
+
+Construct $G_{lat} = (V_{lat}, E_{lat})$ a lattice graph with equal side lengths in $[0,1]^d$, where
+\begin{align*}
+V_{lat} & = P_{lat}(N) := \set{\left(\frac{i_1}{N} - \frac{1}{2N}, \cdots, \frac{i_d}{N} - \frac{1}{2N}\right): i_1, \ldots, i_d \in \set{1, \ldots, N}} \\
+(z,z') & \in E_{lat}~ \text{if and only if $\norm{z - z'} \leq \frac{1}{N}$}
+\end{align*}
+where $z$ and $z' \in P_{lat}(N)$.
+
+Denoting $I = P_{lat}$, we define
+\begin{equation*}
+P_I(x) = \argmin \set{\norm{x - z'}_{\infty}, z' \in P_{lat}(N)}
 \end{equation*}
 
-One such IPM that has not received close attention until surprisingly recently is the \emph{(weighted) Sobolev IPM}.
-For $f \in L^2(\Dset)$ and $\rho$ a density function over $\Dset$, the ($\rho^2$-weighted) Sobolev 1,2 norm of $f$ is given by
+Then, let $C(z) = \set{x \in [0,1]^d: z =  P_I(x)}$ be the collection of cells associated with the mesh $P_{lat}(N)$, noting that $\set{C(z): z \in P_{lat}(N)}$ defines a partition over $[0,1]^d$.
+
+\textcolor{red}{Introduce $\bar{f}$}.
+
+\section{High-Level Proof Strategy for Theorem \ref{thm: weak_convergence_null_hypothesis_T2}}
+
+\begin{enumerate}[(i)]
+	\item \emph{Quantization:} Consider the function class
 	\begin{equation*}
-\norm{f}_{1,2,\rho^2} := 
-\begin{cases}
-\int_{\Dset}\norm{\nabla_x f(x)}^2 \rho^2 dx,~ & \text{$f \in H^1(\Dset)$} \\
-\infty,~ & \text{$f \in L^2(\Dset) \setminus H^1(\Dset)$}
-\end{cases}
+	\Wset_n = \set{\frac{\overline{f}}{\norm{Bf}_2}: f \in \Wset^{1,2}(\Dset,\rho^2)}.
 	\end{equation*}
-where $H^1(\Dset)$ is \textcolor{red}{the Sobolev space of $L^2(\Dset)$ functions with weak derivative $\nabla_x f(x) \in L^2(\Dset)$.}
+	Under assumptions,
+	\begin{equation*}
+	\sup_{\theta: \norm{B\theta}_2 \leq 1} \abs{\lbf^T \theta} - \sup_{\widetilde{f} \in \Wset_n} \abs{\Pbb_n \widetilde{f} - \Qbb_n \widetilde{f}} = o_{\Pbb}(n^{-1/2})
+	\end{equation*}
+	\item \emph{Donsker-convergence of empirical process:} Write 
+	\begin{equation*}
+	\widetilde{f} = \frac{\overline{f}}{\norm{Bf}_2}.
+	\end{equation*}
+	Under \textcolor{red}{assumptions},
+	\begin{equation*}
+	\set{\Gbb_{\Pbb_n}\widetilde{f}: f \in \Wset^{1,2}(\Dset,\rho^2)} \rightsquigarrow G_{\Pbb}
+	\end{equation*}
+	and 
+	\begin{equation*}
+	\set{\Gbb_{\Qbb_n}\widetilde{f}: f \in \Wset^{1,2}(\Dset,\rho^2)} \rightsquigarrow G_{\Qbb},
+	\end{equation*}
+	where \textcolor{red}{$G_{\Pbb}$ is a tight Gaussian process with support on $\Wset^{1,2}(\Dset,\rho^2)$} and for measures $P_n$ and $P$, $\Gbb_{P_n,P} = \sqrt{n}(P_n - P)$ and we suppress notational dependence on $P$ when obvious from context.
+	\item \emph{Continuous mapping:}
+	Under the null hypothesis $H_0: \Pbb = \Qbb$, for any $\widetilde{f} \in \Wset_n$, 
+	\begin{equation*}
+	\sqrt{n}(\Pbb_n \widetilde{f} - \Qbb_n \widetilde{f}) = (\Gbb_{\Pbb_n} - \Gbb_{\Qbb_n})\widetilde{f}
+	\end{equation*}
+	Thus, by i), \textcolor{red}{the independence of $\Pbb_n$ and $\Qbb_n$}, and the continuous mapping theorem,
+	\begin{equation*}
+	\set{\sqrt{n}(\Pbb_n \widetilde{f} - \Qbb_n \widetilde{f}), f \in \Wset^{1,2}(\Dset,\rho^2)} \rightsquigarrow G_{\Pbb} - G'_{\Pbb}
+	\end{equation*}
+	where $G_{\Pbb}$ and $G'_{\Pbb}$ are i.i.d. Gaussian processes.
 
-Consider the unit ball of $\norm{\cdot}_{1,2,\rho^2}$,
+	By the continuous mapping theorem again
 	\begin{equation*}
-\Wset^{1,2}(\Dset,\rho^2) := \set{f: \norm{f}_{1,2,\rho^2} \leq 1}.
+	\sup_{f \in \Wset^{1,2}(\Dset,\rho^2)} \abs{\sqrt{n}(\Pbb_n \widetilde{f} - \Qbb_n \widetilde{f}) } \rightsquigarrow \sup_{f \in \Wset^{1,2}(\Dset,\rho^2)} \abs{(G_{\Pbb} - G'_{\Pbb})f}
 	\end{equation*}
-The weighted Sobolev IPM is simply $\gamma_{\Wset^{1,2}(\Dset,\rho^2)}(\Pbb,\Qbb)$.
+\end{enumerate}
 
-\paragraph{Neighborhood graph.}
+\section{Proof Strategy for \emph{Quantization}}
 
-Let $G_{n,r_n} = (V,E,w)$ denote a weighted, undirected graph constructed from the samples $\Zbf = \set{z_1, \ldots, z_n} = (\Xbf,\Ybf)$ where $V = \set{1, \ldots, n}$, and $w_{uv} = K(z_u, z_v):= k(\frac{\norm{z_u - z_v}}{\epsilon_n}) \geq 0$ for $u,v \in V$, and a particular kernel function $k$.
-Here $(u,v) \in E$ if and only if $w_{uv} > 0$.
+\section{Proof Strategy for ii)}
 
-Motivated by the integral probability metric, our test statistics will be of the form
+We want to exhibit a function class
 \begin{equation*}
-\gamma_{\F_n}(\Pbb_n,\Qbb_n) = \sup_{f_n \in \F_n}\abs{\int_{D} f_n d\Pbb_n - \int_{D} f_n d\Qbb_n}
+\overline{\Wset}_n = \set{f_{n,t}: t \in \Wset^{1,2}(\Dset,\rho^2)}
 \end{equation*}
-where
+where $f_{n,t}: \Dset \to \Reals$ such that:  
+\begin{enumerate}[(a)]
+	\item 
+	There exists a bijection $\phi: \Wset_n \to \overline{\Wset}_n$ with
 	\begin{equation*}
-\Pbb_n := \frac{1}{n_1}\sum_{i = 1}^{n_1} \delta_{x_i}, ~~ \Qbb_n := \frac{1}{n_2}\sum_{i = 1}^{n_2} \delta_{y_i}
+	\sqrt{n}\Pbb_n(f_n - \phi(f_n)) \rightsquigarrow 0.
 	\end{equation*}
-are the empirical distributions of $\Xbf$ and $\Ybf$, respectively, and $\F_n$ is a \textcolor{red}{class of functions $f_n: \Zbf \to \Reals$ exhibiting some regularity with respect to the neighborhood graph $G_{n,r}$.}
-
-\paragraph{Laplacian smoothing and Total Variation denoising.}
-
-For convenience, we number the edges $E  = (e_1, \ldots, e_m)$.
-We denote by $\Bbf \in \Reals^{m \times n}$ the edge incidence matrix of $G_{n,r}$, which for $k$th edge $e_k = (u,v)$ has $k$th row $\Bbf_{k} = (0,\ldots,-w_{uv},\ldots,w_{uv},\ldots,0)$ with a $-w_{uv}$ in the $u$th location, and a $w_{uv}$ in the $v$th location.
-The random (unnormalized) Laplacian matrix is then $\Lbf = \Bbf^T \Bbf$.
-We also introduce a \emph{label vector}, given by $\lbf = (\ell_1, \ldots, \ell_n)$ with
-\begin{equation}
-\label{eqn: label_vector}
-\ell_k = 
-\begin{cases}
-\frac{n}{n_1},~~~ \text{$z_k \in \Xbf$} \\
--\frac{n}{n_2}, ~ \text{$z_k \in \Ybf$}
-\end{cases}
-\end{equation}
-
-Our test statistics $T_1(\lbf; G_{n,r})$ and $T_2(\lbf; G_{n,r})$ are defined as follows:
+	\item 
+	$\overline{\Wset}_n$ is totally bounded.
+	\item
+	For every sequence $\delta_n \downarrow 0$,
+	\begin{equation*}
+	\sup_{\norm{s - t}_{1,2,\rho^2} \leq \delta_n} \Pbb(f_{n,s} - f_{n,t})^2 \to 0
+	\end{equation*}
+	\item
+	There exists a sequence of envelope functions $F_n$ satisfying the Lindeberg condition
 	\begin{align*}
-	T_1(\lbf; G_{n,r}) := \sup_{\fbf \in \Reals^n: \norm{\Bbf \fbf}_1 \leq C_{n,r}} ~ \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_k \\
-	T_2(\lbf; G_{n,r}) := \sup_{\fbf \in \Reals^n: \norm{\Bbf \fbf}_2^2 \leq C_{n,r}} ~ \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_k
+	\Pbb F_n^2 & = \mathcal{O}(1) \\
+	\Pbb F_n^2 \1{F_n > \epsilon \sqrt{n}} & \to 0 \tag{for any $\epsilon > 0$}
 	\end{align*}
-where \textcolor{red}{$C_{n,r} = $} and we write $\fbf = (f_1,\ldots, f_n)$. We note that, as promised, these satisfy the form
-\begin{align*}
-T_1(\lbf; G_{n,r}) & = \sup_{f_n \in TV_n}\abs{\int_{D} f_n d\Pbb_n - \int_{D} f_n d\Qbb_n} \\
-T_2(\lbf; G_{n,r}) & = \sup_{f_n \in \Wset_n}\abs{\int_{D} f_n d\Pbb_n - \int_{D} f_n d\Qbb_n}
-\end{align*}
-where $TV_n = \set{\fbf: \norm{\Bbf \fbf}_1 \leq C_{n,r}}$ and $\Wset_n = \set{\fbf: \norm{\Bbf \fbf}_2 \leq C_{n,r}}$.
-
-\section{Consistency under fixed alternative}
-
-\paragraph{Binomialized data model.}
-For technical reasons, we would like $z_1, \ldots, z_n$ to be independent and identically distributed. We consider the following generative model, which we term the \emph{binomialized data model}:
-
-Fix $n \in \Naturals > 0$, and $n_1 \sim \Bin(n,1/2),~ n_2 = n - n_1$.
-Then, let $x_1, \ldots, x_{n_1} \in \Rd$ be a sequence of i.i.d random points chosen according to $\Pbb$, and $y_1, \ldots, y_{n_2} \in \Rd$ a separate sequence of i.i.d random points chosen according to $\Qbb$, with $x_j \perp y_k$ for all $j,k$.
-Fix $\widetilde{\Zbf} = (\widetilde{z}_1,\ldots,\widetilde{z}_n) := (x_1, \ldots, x_{n_1}, y_1, \ldots, y_{n_2})$.
-Finally, for a permutation $\pi: [n] \to [n]$ chosen uniformly at random among all such permutations, let $\Zbf = (z_1,\ldots,z_n) =  (\widetilde{z}_{\pi(1)}, \ldots, \widetilde{z}_{\pi(n)})$.
-
-The label vector $\lbf$ remains defined as in \eqref{eqn: label_vector} with respect to $\Zbf$.
-Note that now $z_i \overset{i.i.d}{\sim} \frac{\Pbb}{2} + \frac{\Qbb}{2}$, as we desired, with density function $\mu(x) := \frac{p(x) + q(x)}{2}$.
+	\item The bracketing integral
+	\begin{equation*}
+	\int_{0}^{\delta} \sqrt{\log N_{[]}(\epsilon, \overline{W}_n, L_2(\Pbb))}
+	\end{equation*}
+	converges to $0$ for any $\delta_n \downarrow 0$.
+\end{enumerate}
 
 \begin{theorem}
-	\label{thm: LLN_for_t2}
-	Let $d \geq 2$ and let $\Dset \subset \Rd$ be an open, bounded, connected set with Lipschitz boundary. Let $\mu$ satisfy
+	\label{thm: donsker_changing_classes}
+	Let $\F_n = \set{f_{n,t}: t \in T}$ be a class of measurable functions indexed by a totally bounded semimetric space $(T,\rho)$ satisfying
 	\begin{equation*}
-	m \leq \mu(x) \leq M \tag{$\forall x \in D$}
+	\sup_{\rho(s,t) < \delta_n} P (f_{n,s} - f_{n,t})^2 \to 0, ~~ \text{every $\delta_n \downarrow 0$}
 	\end{equation*}
-	for some $0 < m \leq M$. Let $(r_n)$ be a sequence of positive numbers converging to $0$ and satisfying
+	and with envelope function $F_n$ satisfying the Lindeberg condition
 	\begin{align*}
-	& \lim_{n \to \infty} \frac{(\log n)^{3/4}}{n^{1/2}} \frac{1}{r_n} = 0 ~~ \text{if $d = 2$} \\
-	& \lim_{n \to \infty} \frac{(\log n)^{1/d}}{n^{1/d}} \frac{1}{r_n} = 0 ~~ \text{if $d \geq 3$}
+	PF_n^2 & = \mathcal{O}(1) \\
+	PF_n^2 \1 \set{F_n > \epsilon \sqrt{n}} & \to 0, ~~ \textrm{for every $\epsilon > 0$.}
 	\end{align*}
-	Assume the kernel $k$ satisfies conditions: 
-	\begin{align*}
-	& k(0) > 0 ~ \text{and $k$ is continuous at $0$.} \tag{\textbf{K1}} \\
-	& k~ \text{is non-increasing.} \tag{\textbf{K2}} \\
-	& \text{The integral $\int_{0}^{\infty} k(r) r^{d+1} dr$ is finite.} \tag{\textbf{K3}}
-	\end{align*}
-
-	Then with probability one the following statement holds: For $(z_1, \ldots, z_n)$ chosen under the binomialized data model
+	If $J_{[]}(\delta_n, \F_n, L_2(P)) \to 0$ for every $\delta_n \downarrow 0$, then the sequence $\set{\Gbb_n f_{n,t}, t \in T}$ converges in distribution to a tight Gaussian process, provided the sequence of covariance functions
 	\begin{equation*}
-	\lim_{n \to \infty} T_2(\lbf; G_{n,r_n}) = \gamma_{\Wset^{1,2}(\Dset,\mu^2)}(\Pbb,\Qbb).
+	P f_{n,s} f_{n,t} - P f_{n,s} P f_{n,t} 
 	\end{equation*}
+	converges pointwise on $T \times T$. 
 \end{theorem}
 
-\section{Proofs}
-
-Fix $\mu_n = \frac{d\Pbb_n + d\Qbb_n}{2}$. \textcolor{red}{We will show a variational form of convergence of $\mu_n$ to $\mu$.}
-
-\subsection{Gamma convergence of constraint}
-
-\begin{definition}[$TL^2$ convergence]
-	Denote by $\mathfrak{B}(\Dset)$ the Borel $\sigma$-algebra of $\Dset$ and $\mathcal{P}(\Dset)$ the set of all Borel probability measures on $\Dset$.
-Given a Borel map $T: \Dset \to \Dset$, the \emph{push-forward} of $\mu$ by $T$ is given by
-	\begin{equation*}
-	T_{\star}\mu(\Aset) := \mu(T^{-1}(\Aset)), ~~ \Aset \in \mathfrak{B}(\Dset).
-	\end{equation*}
-
-	Given $\widetilde{\mu} \in \mathcal{P}(\Dset)$, we say that $T$ is a \emph{transportation map} between $\mu$ and $\widetilde{\mu}$ if $T_{\star}\mu = \widetilde{\mu}$.
-If for a sequence $(T_n)$ of transportation maps
-	\begin{equation*}
-	\int_{\Dset} \abs{x - T_n(x)}^2 d\mu(x) \to 0,~ \text{as $n \to \infty$}
-	\end{equation*}
-	we refer to the sequence as \emph{stagnating.}
-
-	Take $f \in L^2(\mu)$ and a sequence $(f_n)$ with $f_n \in L^2(\mu_n)$ for $n = 1,2,\ldots$. If there exists a stagnating sequence of transportation maps $(T_n)$ such that
-	\begin{equation*}
-	\int_{\Dset} \abs{ f - T_n(f_n(x)) }^2 d\mu(x) \to 0,~ \text{as $n \to \infty$}
-	\end{equation*}
-	we say that $(f_n)$ converges $TL^2$ to $f$, and write $f_n \overset{TL^2}{\to} f$.
-\end{definition}
-
-We restate Theorem 1.4 of \cite{trillos2018}, changing notation to match the rest of this paper.
-
-\begin{theorem}[Theorem 1.4 of \cite{trillos2018}]
-	Under the setup and conditions of Theorem \ref{thm: LLN_for_t2}, with probability one the following statements hold:
-	\begin{itemize}
-		\item \emph{Liminf inequality:}
-		For all $f \in L^2(\mu)$ and all sequences $(f_n)$ with $f_n \in L^2(\mu_n)$ and $f_n \overset{TL^2}{\to} f$,
-		\begin{equation*}
-		\liminf_{n \to \infty} \frac{1}{n^2 r_n^{d + 2}} \norm{Bf}_2^2 \geq \textcolor{red}{\sigma_{k}} \norm{f}_{1,2,\rho^2}
-		\end{equation*}
-
-		\item \emph{Limsup inequality:} For all $f \in L^2(\mu)$, there exists a sequence $(f_n)$ with $f_n \in L^2(\mu_n)$ and $f_n \overset{TL^2}{\to} f$ such that
-		\begin{equation*}
-		\limsup_{n \to \infty} \frac{1}{n^2 r_n^{d + 2}} \norm{Bf}_2^2 \leq \textcolor{red}{\sigma_{k}} \norm{f}_{1,2,\rho^2}
-		\end{equation*}
-
-		\item \emph{Compactness property:} Every sequence $(f_n)$ with $f_n \in L^2(\mu_n)$ satisfying
-		\begin{equation*}
-		\sup_{n \in \Naturals} \frac{1}{n^2 r_n^{d + 2}} \norm{Bf}_2^2 < \infty
-		\end{equation*}
-		is precompact in $TL^2$, that is, every subsequence of $(f_n)$ has a further subsequence which converges in the $TL^2$-sense to an element of $L^2(\Dset)$.
-	\end{itemize}
-
-\end{theorem}
-
-\subsection{Continuity of risk functional}
-
-\subsection{Proof of Theorem \ref{thm: LLN_for_t2}}
-
-\clearpage
-
-\bibliography{../graph_testing_bibliography}
-\bibliographystyle{plain}
-
 \end{document}

### 2019-02-08 notes/weekly_notes/2.6.19.NotesForWeek.tex
added changes to this weeks notes

@@
 
-\textcolor{red}{Introduce $\bar{f}$}.
+For a given function $f$, the \emph{quantization} $\bar{f}$ over $\set{C(z): z \in P_{lat}(N)}$ is defined by
+\begin{equation*}
+\overline{f}(x_0) = \frac{1}{\nu(C(z_0))} \int_{C(z_0)} f(x) d \nu(x)
+\end{equation*}
+where $x_0 \in C(z_0)$.
 
@@
 
+For any function $f:\Dset \to \Reals$, let $\theta_{f}$ denote the evaluations of $f$ over $\Xbf$, meaning
+\begin{equation*}
+(\theta_f)_i = f(x_i)
+\end{equation*}
+
 \begin{enumerate}[(i)]
 	\item \emph{Quantization:} Consider the function class
-	\begin{equation*}
-	\Wset_n = \set{\frac{\overline{f}}{\norm{Bf}_2}: f \in \Wset^{1,2}(\Dset,\rho^2)}.
-	\end{equation*}
+	\begin{equation}
+	\label{eqn: quantized_function_class}
+	\Wset_n = \set{C_{n,r}\frac{\overline{f}}{\norm{\Bbf \theta_{\bar{f}}}_2}: f \in \Wset^{1,2}(\Dset,\rho^2)}.
+	\end{equation}
 	Under assumptions,
@@
 	\begin{equation*}
-	\widetilde{f} = \frac{\overline{f}}{\norm{Bf}_2}.
+	\widetilde{f} = \frac{\overline{f}}{\norm{\Bbf \theta_{\bar{f}}}_2}.
 	\end{equation*}
@@
 
-\section{Proof Strategy for ii)}
+For a given $\theta \in \Reals^n$, let the quantization of $\theta$ be given by
+\begin{equation*}
+(\bar{\theta})_i = \frac{\sum_{x_j \in C(z_0)} \theta_j}{\abs{C(z_0)}}
+\end{equation*}
+where $x_i \in C(z_0)$, and $\abs{C(z_0)} = \sum_{j = 1}^{n} \1(x_j \in C(z_0))$ is the number of data points in $C(z_0)$.
 
-We want to exhibit a function class
+\begin{lemma}
 	\begin{equation*}
-\overline{\Wset}_n = \set{f_{n,t}: t \in \Wset^{1,2}(\Dset,\rho^2)}
+	\sup_{\theta: \norm{B\theta}_2 \leq C_{n,r}} \abs{\lbf^T \theta} \geq \sup_{\widetilde{f} \in \Wset_n} \abs{\Pbb_n \widetilde{f} - \Qbb_n \widetilde{f}}
 	\end{equation*}
-where $f_{n,t}: \Dset \to \Reals$ such that:  
+\end{lemma}
+\begin{proof}
+	For any $f: \Dset \to \Reals$, $\mathbb{P}_n f - \mathbb{Q}_n f = \lbf^T \theta_f$, and so
+	\begin{equation*}
+	\sup_{\widetilde{f} \in \Wset_n} \abs{\Pbb_n \widetilde{f} - \Qbb_n \widetilde{f}} = \sup_{\theta_{\widetilde{f}}: \widetilde{f} \in \Wset_n} \abs{\lbf^T \theta_{\widetilde{f}}}
+	\end{equation*}
+	Then, for $\widetilde{f} \in \Wset_n$
+	\begin{equation*}
+	\norm{B \theta_{\widetilde{f}}}_2 = \norm{B \theta_{\frac{C_{n,r}\overline{f}} {\norm{B \theta_{\overline{f}}}}_2}} = C_{n,r}.
+	\end{equation*}
+	and so $\set{\theta_{\widetilde{f}}: \widetilde{f} \in \Wset_n} \subset \set{\theta: \norm{\Bbf \theta}_2 \leq C_{n,r}}$.
+\end{proof}
+
+\begin{lemma}
+	For any $\theta \in \Reals^n$
+	\begin{equation*}
+	\lbf^T(\theta - \overline{\theta}) \leq 2 \norm{\Bbf \theta}_{\textcolor{red}{1}}
+	\end{equation*}
+\end{lemma}
+
+\begin{theorem}
+	For any $\theta \in \Reals^n$ such that $\norm{\Bbf \theta}_2 \leq C_{n,r}$, there exists some $f \in \Wset^{1,2}(\Dset,\rho^2)$ such that 
+	\begin{equation*}
+	\abs{\Pbb_n \widetilde{f} - \Qbb_n \widetilde{f} } \geq \abs{\Pbb_n \overline{\theta} - \Qbb_n \overline{\theta} } - o_{\Pbb}(n^{-1/2})
+	\end{equation*}
+\end{theorem}
+\begin{proof}
+	For $\theta \in \Reals^n$, consider the quantized extension of $\theta$, $\overline{f}_{\theta}: \Dset \to \Reals$ defined by
+	\begin{equation*}
+	\overline{f}_{\theta}(x) = (\overline{\theta})_i
+	\end{equation*}
+	for any $x_i$ in the same cell $C(z)$ as $x$. Clearly,
+	\begin{equation*}
+	\Pbb_n \overline{f}_{\theta} - \Qbb_n \overline{f}_{\theta} = \Pbb_n \overline{\theta} - \Qbb_n \overline{\theta} 
+	\end{equation*}
+	Now, we have to show that either $\overline{f}_{\theta} \in \Wset_n$, or that it can be well-approximated by some $\widetilde{f} \in \Wset_n$. 
+\end{proof}
+
+\section{Proof strategy for \emph{Donsker-convergence of empirical processes}:}
+
+Consider the function class $\Wset_n$ given by \eqref{eqn: quantized_function_class}. We wish to show
+
 \begin{enumerate}[(a)]
 	\item 
-	There exists a bijection $\phi: \Wset_n \to \overline{\Wset}_n$ with
-	\begin{equation*}
-	\sqrt{n}\Pbb_n(f_n - \phi(f_n)) \rightsquigarrow 0.
-	\end{equation*}
-	\item 
-	$\overline{\Wset}_n$ is totally bounded.
+	$\Wset_n$ is totally bounded.
 	\item
@@
 	\begin{equation*}
-	\sup_{\norm{s - t}_{1,2,\rho^2} \leq \delta_n} \Pbb(f_{n,s} - f_{n,t})^2 \to 0
+	\sup_{\norm{f - g}_{1,2,\rho^2} \leq \delta_n} \Pbb(\widetilde{f} - \widetilde{g})^2 \to 0
 	\end{equation*}
+	where $f$ and $g$ are arbitrary elements of $\Wset^{1,2}(\Dset,\rho^2)$.
 	\item
-	There exists a sequence of envelope functions $F_n$ satisfying the Lindeberg condition
+	There exists a sequence of envelope functions $F_n$ on $\Wset_n$ satisfying the Lindeberg condition
 	\begin{align*}
@@
 	\begin{equation*}
-	\int_{0}^{\delta} \sqrt{\log N_{[]}(\epsilon, \overline{W}_n, L_2(\Pbb))}
+	\int_{0}^{\delta} \sqrt{\log N_{[]}(\epsilon, \Wset_n, L_2(\Pbb))}
 	\end{equation*}
@@
 
-\begin{theorem}
+Then, the desired results holds by application of Theorem \ref{thm: donsker_changing_classes}.
+
+\begin{theorem}[Donsker theorem for changing function classes]
 	\label{thm: donsker_changing_classes}
+
 	Let $\F_n = \set{f_{n,t}: t \in T}$ be a class of measurable functions indexed by a totally bounded semimetric space $(T,\rho)$ satisfying

### 2019-02-15 notes/weekly_notes/2.14.19.NotesForWeek.tex
added about half of weekly notes

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbb{I}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Matrices
\newcommand{\Xbf}{\mathbf{X}}
\newcommand{\Ybf}{\mathbf{Y}}
\newcommand{\Zbf}{\mathbf{Z}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 2/8/19 - 2/15/19}
\author{Alden Green}
\date{\today}
\maketitle

Consider $\Xbf = (X_1, \ldots, X_n) \sim \Pbb$ and $\Ybf = (Y_1, \ldots, Y_m) \sim \Qbb$ independently generated random variables. We will assume $\Pbb$ and $\Qbb$ are supported on $\D \subset \Reals^d$ and are absolutely continuous with density functions $f$ and $g$, respectively. Our statistical goal is hypothesis testing: that is, we wish to construct a test function $\phi$ which differentiates between
\begin{equation*}
\mathbb{H}_0: f = g \text{ and } \mathbb{H}_1: f \neq g.
\end{equation*}

For a given function class $\Hclass$, some $\epsilon > 0$, and test function $\phi$ a Borel measurable function of the data with range $\set{0,1}$, we evaluate the quality of the test using \emph{worst-case risk}
\begin{equation*}
R_{\epsilon}^{m,n}(\phi; \Hclass) = \sup_{f \in \Hclass} \Ebb_{f,f}^{(m,n)}(\phi) + \sup_{ \substack{f,g \in \Hclass \\ \delta(f,g) \geq \epsilon } } \Ebb_{f,g}^{m,n}(1 - \phi)
\end{equation*}

\paragraph{Total variation test.}

Let $\Zbf = (X_1, \ldots, X_n, Y_1, \ldots, Y_m)$, and let $t = n + m$. As in \citep{padilla2018}, define the \emph{$K$-NN graph} $G_K = (V,E_K)$ to have vertex set $V = \set{1, \ldots, t}$ and edge set $E_K$ which contains the pair $(i,j)$ if and only if $x_i$ is among the $K$-nearest neighbors (with respect to Euclidean distance) of $x_j$, or vice versa. Let $D_K$ denote the incidence matrix of $G_K$.

Define the \emph{kNN-total variation} test statistic to be
\begin{equation}
\label{eqn: knn_tv_test_statistic}
T_{TV} = \sup_{\substack{\theta \in \Reals^{t}: \\ \norm{D_{k} \theta}_1 \leq C_{n,k}} } \left(\sum_{i = 1}^{n} \theta_i - \sum_{j = n+1}^{t} \theta_j\right)
\end{equation}

Hereafter, take $\D = [0,1]^d$, and consider
\begin{equation*}
\Hclass_{lip}(L) = \set{f: [0,1]^d \to \Reals^{+}: \int_{\D}f = 1, f~ \text{$L$-piecewise Lipschitz, bounded above and below} }
\end{equation*}
where
\begin{definition}[Piecewise Lipschitz]
	\label{def: piecewise_lipschitz}
	A function $f$ is $L$-\emph{piecewise lipschitz} over $[0,1]^d$ if there exists a set $\Sset \subset [0,1]^d$ such that
	\begin{enumerate}[(a)]
		\item $\nu(\Sset) = 0$
		\item There exist $C_{\Sset}, \epsilon_0$ such that $\mu \Bigl( \bigl(\Sset_{\epsilon} \cup (\partial \D)_{\epsilon} \bigr) \cap [0,1]^d \Bigr) \leq C_{\Sset} \epsilon$ for all $0 < \epsilon \leq \epsilon_0$.
		\item For any $z, z'$ in the same connected component of $[0,1]^d \setminus \bigl(\Sset_{\epsilon} \cup (\partial \D)_{\epsilon}\bigr)$,
		\begin{equation*}
		\abs{g(z) - g(z')}_2 \leq L\norm{z - z'}_2
		\end{equation*}
	\end{enumerate}
\end{definition}
and
\begin{definition}[Bounded above and below]
	\label{def: bounded}
	A function $f: \D \to \Reals$ is \emph{bounded above and below} if there exists $p_{\min}, p_{\max}$ such that
	\begin{equation*}
	0 < p_{\min} < f(x) < p_{\max} < \infty \tag{$\forall x \in \D$}
	\end{equation*}
\end{definition}

\begin{conjecture}
	For $\tau = ???$ and $K \asymp \log^{1 + 2r}(n)$ for some $r \geq 0$, the test $\phi_{TV} = \set{T_{TV} \geq \tau}$ has worst-case risk
	\begin{equation*}
	R_{\epsilon}^{(m,m)}(\Hclass_{lip}(L)) \leq c_1/a^2
	\end{equation*}
	whenever $\epsilon \geq c_2 a \log^{\alpha}m m^{-1/d}$ where $\alpha = 3r + 5/2 + (2r + 1)/d$ and $c_1$ and $c_2$ are constants which depend only on $(d,L)$.
\end{conjecture}
\begin{proof}
	Write
	\begin{equation*}
	\sum_{i = 1}^{m} \theta_i - \sum_{j = m + 1}^{2m} \theta_j = \dotp{\theta}{1_X - 1_Y}
	\end{equation*}
	where
	\begin{equation*}
	1_X = 
	\begin{cases}
	1,~ i = 1, \ldots,m \\
	0,~ \text{otherwise}
	\end{cases}
	1_Y = 
	\begin{cases}
	1,~ i = j + 1, \ldots,2m \\
	0,~ \text{otherwise}
	\end{cases}
	\end{equation*}
	Let
	\begin{equation*}
	\widehat{\theta} \in \argmax_{\theta \in \Reals^{2m}} \set{\dotp{\theta}{1_X - 1_Y}: \norm{D_K \theta}_1 \leq C_{n,k}}
	\end{equation*}
	satisfy $T_{TV} = \dotp{\widehat{\theta}}{1_X - 1_Y}$. For $i = 1, \ldots, 2m$, introduce $\theta^{\star}$ defined by
	\begin{equation*}
	(\theta^{\star})_i = \left(\frac{f(z_i) - g(z_i)}{f(z_i) + g(z_i)} \right)	\end{equation*}

	We have
	\begin{equation*}
	T_{TV} = \dotp{\widehat{\theta} - \theta^{\star}}{1_X - 1_Y} + \dotp{\theta^{\star}}{1_X - 1_Y}
	\end{equation*}

	We bound the first term, and begin by employing a basic inequality argument. Let
	\begin{align*}
	\widetilde{\theta} = \frac{\theta^{\star}}{\norm{D\theta^{\star}}_1}.
	\end{align*}
	Since $\widetilde{\theta}$ is feasible and $\widehat{\theta}$ optimal for \eqref{eqn: knn_tv_test_statistic}, we have
	\begin{equation*}
	\dotp{\widehat{\theta}}{1_X - 1_Y} \geq \dotp{\widetilde{\theta}}{1_X - 1_Y}
	\end{equation*}
	which, letting $w = (1_X - 1_Y) - \theta^{\star}$, means
	\begin{equation}
	\label{eqn: basic_inequality_1}
	\dotp{\widehat{\theta}}{\theta^{\star}} \geq \dotp{\widetilde{\theta}}{\theta^{\star}} - \dotp{\widehat{\theta} - \widetilde{\theta}}{w}
	\end{equation}

	As a result, we have
	\begin{align*}
	\abs{\dotp{\widehat{\theta} - \theta^{\star}}{1_X - 1_Y}} & \leq \abs{\dotp{\widehat{\theta} - \theta^{\star}}{\theta^{\star}} } + \abs{\dotp{\widehat{\theta} - \theta^{\star}}{w} } \\
	& \overset{(i)}{\leq} \abs{\dotp{\widetilde{\theta} - \theta^{\star}}{\theta^{\star}} } + \abs{\dotp{\widehat{\theta} - \widetilde{\theta}}{w} } + \abs{\dotp{\widehat{\theta} - \theta^{\star}}{w} } \\
	& \leq \abs{\dotp{\widetilde{\theta} - \theta^{\star}}{\theta^{\star}} } + 2\abs{\dotp{\widehat{\theta} - \widetilde{\theta}}{w} } + \abs{\dotp{\widetilde{\theta} - \theta^{\star}}{w}}
	\end{align*}
	where $(i)$ follows from \eqref{eqn: basic_inequality_1}.
\end{proof}

\clearpage

\bibliography{../../graph_testing_bibliography}
\bibliographystyle{plain}

\end{document}

### 2019-02-15 notes/reference_notes/ariascastro17.tex
added updated version of this week's notes and notes on references

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbb{I}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Matrices
\newcommand{\Xbf}{\mathbf{X}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}

\title{Notes on 'Remember the Curse of Dimensionality: The Case of Goodness-of-Fit Testing in Arbitrary Dimension'}
\author{Alden Green}
\date{\today}
\maketitle

Define the \emph{worst-case risk} of a test $\phi$ (Borel-measurable function to $[0,1]$) to be
\begin{equation*}
R_{\epsilon}^{(m)}(\phi; f_0; \mathcal{H}) = \Ebb_{f_0}^{(m)} \phi + \sup_{\substack{f \in \Hclass \\ \delta(f,f_0) \leq \epsilon}} \set{\Ebb_f^{(m)} ( 1 - \phi)}
\end{equation*}
where $\delta(f,g) = \int (f-g)^2$.

The \emph{minimax risk} is then
\begin{equation*}
R_{\epsilon}^{(m)}(f_0; \mathcal{H}) = \inf_{\phi} R_{\epsilon}^{(m)}(\phi; f_0; \mathcal{H}).
\end{equation*}

\section{One-sample goodness of fit problem for Holder class}

Let $\Hclass_s^d(L)$ be the \emph{Holder class} of functions $f: [0,1]^d \to \Reals$ such that
\begin{equation*}
\abs{f^{\floor{s}}(x) - f^{\floor{s}}(y)} \leq L \norm{x - y}^{s - \floor{s}} \tag{for all $x,y \in [0,1]^d$}
\end{equation*}
and additionally
\begin{equation*}
\norm{f^{(s')}}_{\infty} \leq L,~~ \forall s' \in \set{0,\cdots,\floor{s}}
\end{equation*}

\begin{theorem}[One-sample lower bound over Holder densities]
	\label{thm: one_sample_Holder_lower_bound}
	For the one-sample problem
	\begin{equation*}
	f_0 = \text{uniform distribution over} ~[0,1]^d~ \text{known}
	\end{equation*}
	under known Holder regularity, there is a constant $c > 0$ depending only on $(s,d,L)$ such that
	\begin{equation*}
	\R_{\epsilon}^{(m)}(f_0; \Hclass_s^d(L)) \geq 1/2~ \text{for all $\epsilon \leq cm^{-\frac{2s}{4s + d}}$}
	\end{equation*}
\end{theorem}

\subsection{One-sample chi-squared test}

Let the \emph{one sample chi-squared statistic} be given by
\begin{equation*}
\Gamma_{\kappa}^{one} = \sum_{k \in [\kappa]^d} \abs{M_{k,\kappa} - m\kappa^{-d}}^2
\end{equation*}
where
\begin{equation*}
M_{k,\kappa} = \# \set{x_i: x_i \in \left(\frac{k-1}{\kappa}, \frac{k}{\kappa}\right]}
\end{equation*}

For now, we treat $s$ as known, and set
\begin{equation*}
\kappa = \kappa(s,d) := \floor{m^{\frac{2}{4s+d}}}
\end{equation*}

\begin{theorem}[One-sample chi-squared test]
	\label{thm:one_sample_chi_squared_test}
	In the one-sample problem under known Holder regularity, consider the chi-squared test $\phi_{\kappa,\tau} = \1\left\{\Gamma_{\kappa}^{one} > \tau \right\}$. There are constants $c_1$ depending on only $(s,L)$ and $c_2$ depending on only $(s,d,L)$ such that for $\tau = m + a m \kappa^{-d/2}$ with $a \geq 1$
	\begin{equation*}
	R_{\epsilon}^{(m)}(\phi_{\kappa,\tau}; f_0, \mathcal{H}_s^d(L)) \leq c_1/a^2
	\end{equation*}
	for any $\epsilon \geq c_2 a m^{-\frac{2s}{4s+d}}$.
\end{theorem}

\section{Two-Sample Testing}
In the two sample case, we observe data $x_1, \ldots, x_n$ and $y_1, \ldots, y_m$. For a given test (Borel measurable function of the data) $\phi$, define the worst case risk to be
\begin{equation*}
R_{\epsilon}^{(m,n)}(\phi; \mathcal{H}) = \sup_{f \in \Hclass} \Ebb_{f,f}(\phi) + \sup_{ \substack{f,g \in \Hclass \\ \delta(f,g) \geq \epsilon^2} } \Ebb_{f,g}(1 - \phi)
\end{equation*}
and the minimax risk to be
\begin{equation*}
R_{\epsilon}^{(m,n)}(\Hclass) = \inf_{\phi} R_{\epsilon}^{(m,n)}(\phi; \mathcal{H}).
\end{equation*}

It is intuitively clear that the two-sample problem is at least as hard as the one-sample problem.

\begin{lemma}[Two-sample is harder than one-sample.]
	\label{lem: two_sample_harder_one_sample}
	For any class $\Hclass$, pseudo-metric $\delta$, $\epsilon > 0$, and $f_0 \in \Hclass$,
	\begin{equation*}
	R_{\epsilon}^{(m)}(f_0; \Hclass) \leq R_{\epsilon}^{(m,n)}(\Hclass)
	\end{equation*}
\end{lemma}

\begin{theorem}[Two-sample lower bound over Holder densities]
	\label{thm: two_sample_Holder_lower_bound}
	For the two-sample problem under known Holder regularity, there exists constant $c$ depending only on $(s,d,L)$ such that
	\begin{equation*}
	R_{\epsilon}^{(m,n)}(\Hclass_s^d(L)) \geq 1/2
	\end{equation*}
	for any $\epsilon \leq c(m \wedge n)^{-\frac{2s}{4s + d}}$. 
\end{theorem}

\subsection{Two-sample chi-squared test.}
Define the \emph{two-sample chi-squared} test statistic
\begin{equation*}
\Gamma_{\kappa} = \sum_{k \in [\kappa]^d} (M_{k, \kappa} - N_{k, \kappa})^2
\end{equation*}
where
\begin{equation*}
M_{k,\kappa} = \# \set{x_i: x_i \in \left(\frac{k-1}{\kappa}, \frac{k}{\kappa}\right]},~~ N_{k,\kappa} = \# \set{y_i: y_i \in \left(\frac{k-1}{\kappa}, \frac{k}{\kappa}\right]}
\end{equation*}
are bin counts, and for simplicity we let $m = n$ so no normalization is needed.

\begin{theorem}[Two-sample chi-squared test.]
	\label{thm: two_sample_chi_squared_test}
	For the two-sample testing problem under known Holder regularity, let $\phi_{k, \kappa} = \mathbb{I}(\Gamma_{\kappa} > \tau)$. There are constants $c_1$ depending only on $L$, and $c_2$ depending only on $(s,d,L)$ such that for $\tau = 2m + a m \kappa^{-d/2}$
	\begin{equation*}
	R_{\epsilon}^{(m,m)}(\phi_{k, \kappa}) \leq c_1/a^2
	\end{equation*}
	whenever $\epsilon \geq a c_2 m^{-\frac{2s}{4s + d}}$. 
\end{theorem}

\section{Proofs}

\begin{proof}[Proof of Theorem \ref{thm: one_sample_Holder_lower_bound}]
	Let $h: \Rd \to \Reals$ be infinitely differentiable with support on $[0,1]^d$ such that $\int h = 0$, $\int h^2 = 1$. Let $\kappa \geq 1$ be an integer, and for $j \in \mathbb{Z}^d$, define
	\begin{equation*}
	h_{j,\kappa}(x) = \kappa^{d/2} h(\kappa x - j + 1)
	\end{equation*}
	supported on $[\frac{(j - 1)}{\kappa}, \frac{j}{\kappa}]$. (Note that $\norm{h_{j,\kappa}} = 1$)

	For $\eta = (\eta_1, \ldots, \eta_{\kappa^d}) \in \set{-1, +1}^{\kappa^d}$ and $\rho > 0$ to be defined later, define
	\begin{equation*}
	f_\eta = f_0 + \rho \sum_{j \in [\kappa]^d} \eta_j h_{j,\kappa}(x)
	\end{equation*}
	and note that since $\int h_{j,\kappa} = 0$, $\int f_{\eta} = 1$. Additionally, because the $h_{j,\kappa}$'s have disjoint support,
	\begin{itemize}
		\item 
		If $\rho \kappa^{d/2} \norm{h}_{\infty} \leq 1$,
		\begin{equation*}
		f_{\eta} \geq 0
		\end{equation*}

		\item
		For $C := 4 \norm{h^{(\floor{s})}}_{\infty} \vee 2 \norm{h^{(\floor{s } + 1)}}_{\infty}$, if $\rho \kappa^{d/2 + s} C \leq L$, then
		\begin{equation*}
		h_{j,\kappa}^{(\floor{s})} \in \Hclass_s^d(L)
		\end{equation*}
	\end{itemize}

	To see the second point, for arbitrary $x,y \in [0,1]^d$ let $x \in [\frac{k-1}{\kappa}, \frac{k}{\kappa}]$, $y \in [\frac{l-1}{\kappa}, \frac{l}{\kappa}]$. Then
	\begin{align*}
	\abs{f_{\eta}^{(\floor{s})}(x) - f_{\eta}^{(\floor{s})}(y)} & \leq 2 \rho \kappa^{d/2 + \floor{s}} \bigl(\abs{h^{(\floor{s})}(\kappa x - k  + 1) - h^{(\floor{s})}(\kappa y - k  + 1)} + \\
	& \abs{h^{(\floor{s})}(\kappa x - l  + 1) - h^{(\floor{s})}(\kappa y - k + 1)} \bigr) \\
	& \leq 2 \rho \kappa^{d/2 + \floor{s}} \left( 2 \kappa \norm{h^{(\floor{s} + 1)} }_{\infty} \norm{x - y} \wedge 4 \norm{h^{(\floor{s})}}_{\infty} \right) \\
	& \leq  2 \rho \kappa^{d/2 + \floor{s}} \left( \left[2 \norm{h^{(\floor{s} + 1)}}_{\infty} \vee 4 \norm{h^{(\floor{s})}}_{\infty} \right] \left[1 \wedge \kappa \norm{x - y}\right]  \right) \\
	& \leq L \norm{x - y}^{s - \floor{s}}
	\end{align*}
	where the step follows from $(1 \wedge u) \leq u^a$ for all $u > 0, 0 < a \leq 1$.

	Take $\rho$ small enough to satisfy the above conditions, let $\epsilon = \rho \kappa^{d/2}$. Note that
	\begin{equation*}
	\norm{f_0 - f_{\eta}}_2^2 \leq \rho^2 \sum_{j \in [\kappa]^d} \norm{h_{j,\kappa}}_2^2 = \rho^2 \kappa^d = \epsilon^2.
	\end{equation*}

	The minimax risk is lower bounded by the Bayes risk; in this case, consider  the uniform prior distribution over $\set{f_{\eta}: \eta \in \set{-1,1}^{[\kappa]^d}}$. The Bayes risk is achieved by the likelihood ratio test $\set{W > 1}$, where
	\begin{equation*}
	W = \frac{1}{2^{\kappa^d}} \sum_{\eta \in \set{-1,1}^{\kappa^d}} \prod_{i = 1}^{m} f_{\eta}(x_i)
	\end{equation*}
	and its known that the risk of the likelihood ratio test is lower bounded by $1 - \sqrt{\frac{1}{2} \Var_{f_0}(W)}$.\footnote{Compare the $\chi^2$-divergence and $TV$-distance, and use the lower bound on risk $R \geq 1/2 - 1/2\norm{P_0 - P_1}_{TV}$.}

	We can compute $\Var_{f_0}(W) \leq \exp \{(\rho^2 m)^2\kappa_d \}$ for $rho^2 m \leq 1$. Choosing $\kappa = \floor{m^{2/(4s + d)}}$ and $\rho = c m^{-(2s + d)/(4s + d)}$, we have that the upper bound on $\Var_{f_0}(W)$ is $1$, and so the minimax risk is at least $1/2$. $\epsilon = \kappa^{d/2} \rho = cm^{\frac{-2s}{4s + d}}$. It can be verified that the choice of $\rho$ satisfies the necessary conditions imposed above. 
\end{proof}

\subsection{Proof of Theorem \ref{thm: two_sample_chi_squared_test}}

\paragraph{Test error for chi-squared test over discrete distributions}

Let $p$ and $q$ be discrete distributions over $\mathcal{K}$, and let
\begin{equation*}
T = \sum_{k \in \mathcal{K}} (M_k - N_k)^2
\end{equation*}
where
\begin{equation*}
M_k = \#\{A_i = k\},~ N_k = \#\{B_j = k\}
\end{equation*}
for $A_1, \ldots, A_m \sim p, B_1, \ldots, B_m \sim q$ independent.

\begin{lemma}[Moment bounds for $T$]
	We have
	\label{lem: moment_bounds_for_T}
	\begin{align*}
	\Ebb(T) & = 2m + m^2 \langle (p-q)^2  \rangle - m(\langle p^2 \rangle + \langle q^2 \rangle) \\
	\Var(T) & = 2m^2 \langle (p + q)^2 \rangle + 4m^3(\langle (p + q)(p - q)^2 \rangle + 2 \langle pq \rangle \langle (p-q^2)\rangle)
	\end{align*}
\end{lemma}

\begin{corollary}
	Consider testing within the class of probability distributions $r$ on $\mathcal{K}$ such that $\norm{r}_{\infty} \leq \eta$ for some $\eta > 0$. There are universal constants $\nu_1$ and $\nu_2$ such that for any $a > 0$, the test with rejection region
	\begin{equation*}
	\set{T - 2m \geq a m \sqrt{\eta}}
	\end{equation*}
	has size at most $\nu_1/a^2$ and power at least $1 - \nu_1/a^2$ against alternatives satisfying
	\begin{equation*}
	\norm{p - q}^2 \geq \nu_2(a \sqrt{\eta} \vee a^2 \eta \vee \eta) / m
	\end{equation*} 
\end{corollary}

\paragraph{Approximation Error.}
\begin{lemma}[Approximation error of binning]
	\label{lem: approximation_error_of_binning}
	For a continuous function $h: [0,1]^d \to \Reals$ and an integer $\kappa \geq 2$, define
	\begin{equation*}
	W_{\kappa}[h] = \sum_{k \in [\kappa^d]} \kappa^d \int_{H_k} h(x) dx \1(H_k)
	\end{equation*}
	Then there are constants $b_1, b_2 > 0$ depending only on $(s,d,L)$ such that
	\begin{equation*}
	\norm{W_k[h]}_2 \geq b_1 \norm{h}_2 - b_2 \kappa^{-s},~ \forall h \in H_s^d(L) 
	\end{equation*}
\end{lemma}

\begin{proof}[Proof of Lemma \ref{lem: approximation_error_of_binning}]

	\begin{lemma}[Taylor expansion]
		\label{lem: taylor_expansion}
		Fix any $h \in \Hclass_s^d(L)$ and any $x_0 \in [0,1]^d$. Let $u$ denote the $\floor{s}$-th order Taylor expansion of $h$ around $x_0$. Then, there is a constant $L'$ depending only on $(s,d,L)$ such that
		\begin{equation*}
		\abs{h(x) - u(x)} \leq L' \norm{x - x_0}^s \tag{$\forall u \in [0,1]^d$}
		\end{equation*}
	\end{lemma}

	Fix $h \in \Hclass_s^d(L)$, and let $u_j$ be the $\floor{s}$-th order Taylor expansion around $(j - 1)r/\kappa$, for some $j = (j_1,\ldots,j_d)$, $r \geq 1$ is an integer. Then, define
	\begin{equation*}
	u = \sum_{j} u_j \1_{\widetilde{\Hclass}_j},~ \widetilde{\Hclass}_j = \prod_{l = 1}^{m} \widetilde{\Hclass}_{j_l},~ \widetilde{\Hclass}_{j_l} = \left( \frac{(j_l - 1)r}{\kappa}, \frac{j_l r}{\kappa} \right]
	\end{equation*}
	Then,
	\begin{equation*}
	\abs{u(x) - h(x)} \leq L'\left(\frac{2\sqrt{d} r}{\kappa}\right)^s =: c_1 \kappa^{-s}
	\end{equation*}
	and as a result
	\begin{align*}
	\norm{W_{\kappa}[h]}_2 & \geq \norm{W_{\kappa}[u]}_2 - \norm{W_{\kappa}[h] - W_{\kappa}[u]}_2 \\
	& \geq \norm{W_{\kappa}[u]}_2 - \norm{h - u}_2 \\
	& \geq \norm{W_{\kappa}[u]}_2 - c_1 \kappa^{-s}
	\end{align*}

	\begin{lemma}[Averaging of polynomial preserves norm]
		\label{lem: averaging_of_polynomial_preserves_norm}
		Let $\Pclass_m^d$ denote the class of polynomials on $\Reals^d$ of degree at most $m$. For a given partition $\mathcal{Q} = (Q_i)$, define
		\begin{equation*}
		W_{\mathcal{Q}}[v] = \sum_{Q_i \in \mathcal{Q}} \frac{1}{\abs{Q_i}} \left(\int_{Q_j} v(x) dx\right) \1_{Q_i}.
		\end{equation*}
		Then there exists constant $c_2 > 0$ depending only on $(d,m)$ such that, when $\max_{j} \diam(Q_j) \leq c_2$,
		\begin{equation*}
		\norm{W_{\mathcal{Q}}[v]} \geq c_2 \norm{v}_2,~ \text{for all $v \in \Pclass_m^d$}.
		\end{equation*}
	\end{lemma}
\end{proof}

\begin{proof}[Proof of Theorem \ref{thm: two_sample_chi_squared_test}]

Define
\begin{equation*}
H_k = \left(\frac{k-1}{\kappa}, \frac{k}{\kappa}\right],~ p_k = \int_{H_k} f(x) dx,~ q_k = \int_{H_k} g(x) dx
\end{equation*}
and let $p = (p_k)_{k \in [\kappa]^d}$, $q = (q_k)_{k \in [\kappa]^d}$.

Application of Lemma \ref{lem: approximation_error_of_binning} to $f - g$ yields
\begin{equation*}
\norm{p - q}^2 \geq \kappa^{-d} \left( b_1^* \norm{f - g}_2 - b_2^* \kappa^{-s} \right)^2.
\end{equation*}

Recall that $\norm{g - f}_2 \geq \epsilon \geq c_1 a m^{-\frac{2s}{4s + d}}$. Since $\kappa^{-s} = m^{-\frac{2s}{4s + d}}$, an appropriate choice of $c_1$, independent of $\kappa$, yields
\begin{equation*}
\norm{p - q}^2 \geq \kappa^{-d} \epsilon^2
\end{equation*}

\end{proof}

\section{Relevant Citations}

\begin{enumerate}
	\item A distribution free version of the smirnov two-sample test in the p-variate case (Bickel 69)
	\item Optimal kernel choice for large-scale two-sample tests. (Gretton 12)
	\item Permutation tests for equality of distributions in high-dimensional settings (Hall 02)
\end{enumerate}

\end{document}

### 2019-02-15 notes/reference_notes/padilla18.tex
added updated version of this week's notes and notes on references

@@
 \newcommand{\dotp}[2]{\langle #1, #2 \rangle}
+\newcommand{\1}{\mathbb{I}}
 
@@
 	\end{equation*}
-	\item The function $g_0 := f_0 \circ h^{-1}$ has bounded variation, meaning
-	\begin{equation*}
-	\abs{g_0}_{BV(\Omega)} := \sup \set{\int_{\Omega} g_0(x) \dive(g)(x)dx; g \in C_c^1(\Omega, \Rd), \norm{g}_{\infty} \leq 1} < \infty
-	\end{equation*}
-	where $\mathfrak{B}$ is the Borel $\sigma$-algebra of $\Omega = [0,1]^d$, $C_c^1(\Omega, \Rd)$ is the set of $\mathfrak{B}$-measurable, continuously differentiable functions into $\Rd$, and
-	\begin{equation*}
-	\norm{g}_{\infty} := \norm{\left(\sum_{i=1}^{d}g_i^2\right)^{1/2}}_{L_{\infty}(\Omega)}.
-	\end{equation*}
 	\item $g_0$ is piecewise Lipschitz\footnote{Technically, the requirement is slightly weaker than piecewise Lipschitz.}, meaning there exists a set $\Sset \subset (0,1)^d$ such that
@@
 
-\section{Proof of Theorem \ref{thm: padilla18_theorem_1}}
+\section{Proofs}
 
@@
 \begin{lemma}
+	\label{lem: quantization_error}
 	Fix $N$ to satisfy \eqref{eqn: N_lower_bound}, and let us assume that the event $\Omega$ from Lemma \ref{lem: mesh_counts} holds.
@@
 \end{lemma}
+\begin{proof}
+	Clearly
+	\begin{equation*}
+	\dotp{\epsilon^T}{\theta - \theta_{I}} \leq \norm{\epsilon}_{\infty} \cdot \norm{\theta - \theta_I}_1
+	\end{equation*}
+	Then, for every $i = 1, \ldots, n$, the event $\Omega$ implies that there exists a $j \in [n]$ such that
+	\begin{equation*}
+	(\theta_I)_i = \theta_j,~~ (x_i,x_j) \in E_{G_K}
+	\end{equation*}
+	and therefore
+	\begin{equation*}
+	\norm{\theta - \theta_I}_1 \leq \norm{\nabla_{G_K} \theta}_1
+	\end{equation*}
+\end{proof}
+
+\subsection{Bounding empirical process}
+
+\begin{lemma}
+	Conditional on the event $\Omega$, we have that
+	\begin{equation*}
+	\dotp{\epsilon}{\widehat{\theta}_I - \widehat{\theta}_I^{\star}} \leq \max_{u \in I} \sqrt{\abs{C(u)}} \left( \norm{\Pi \widetilde{\epsilon}}_2 \norm{\widehat{\theta} - \theta^{\star}}_2  + \norm{(D^{\dagger})^T \widetilde{\epsilon}}_{\infty} \left[ \norm{\nabla_{G_K} \widehat{\theta} }_1 + \norm{\nabla_{G_K} \theta^{\star} }_1  \right]\right)
+	\end{equation*}
+	where $\widetilde{\epsilon}$ is an independent, mean-zero vector of subgaussian random variables, $\abs{C(u)} := \sum_{i \in [n]} \1(x_i \in C(u))$, and $D^{\dagger}$ is the pseudoinverse of the incidence matrix $D$ of $G_{lat}$.
+\end{lemma}
+\begin{proof}
+	Writing
+	\begin{equation*}
+	\widetilde{\epsilon}_l = \left[\max_{u \in I} \abs{C(u)} \right]^{-1/2} \sum_{x_j \in I_l} \epsilon_j
+	\end{equation*}
+	we have
+	\begin{align*}
+	\dotp{\epsilon}{\widehat{\theta}_I - \theta_I^{\star}} = \left[\max_{u \in I} \abs{C(u)}\right]^{1/2} \dotp{\widetilde{\epsilon}}{\widehat{\theta}^{I} - \theta^{\star,I}}
+	\end{align*}
+	Now, divide up $\widetilde{\epsilon}$ into 
+	\begin{equation*}
+	\widetilde{\epsilon} = P_{1}(\widetilde{\epsilon}) + P_{1^{\perp}}(\widetilde{\epsilon})
+	\end{equation*}
+	where $P_1$ is the projection onto the span of $\mathbf{1}$ the constant vector, and $P_{1^{\perp}}$ the projection onto the space orthogonal to $\mathbf{1}$.
+Note that $P_{1^{\perp}}(x) = (D^{\dagger} D)^T x$.
+Then, we have
+	\begin{align*}
+	\dotp{P_{1^{\perp}}(\widetilde{\epsilon})}{\widehat{\theta}^{I} - \theta^{\star,I}} & = \dotp{(D^{\dagger} D)^T \widetilde{\epsilon}}{\widehat{\theta}^{I} - \theta^{\star,I}} \\
+	& = \dotp{(D^{\dagger})^T \widetilde{\epsilon}}{D(\widehat{\theta}^{I} - \theta^{\star,I})} \\
+	& \leq \norm{(D^{\dagger})^T \widetilde{\epsilon}}_{\infty} \norm{ D(\widehat{\theta}^{I} - \theta^{\star,I})  }_1
+	& \leq \norm{(D^{\dagger})^T \widetilde{\epsilon}}_{\infty} \left[ \norm{\nabla_{G_K} \widehat{\theta} }_1 + \norm{\nabla_{G_K} \theta^{\star} }_1 \right]
+	\end{align*}
+	where the last inequality follows from the triangle inequality and Lemma \ref{lem: quantization_error}.
+
+	On the other hand,
+	\begin{equation*}
+	\dotp{P_{1}(\widetilde{\epsilon})}{\widehat{\theta}^{I} - \theta^{\star,I}} \leq \norm{P_{1}(\widetilde{\epsilon})}_2 \norm{\widehat{\theta}^{I} - \theta^{\star,I}}_2
+	\end{equation*}
+	and so the desired result follows.
+\end{proof}
+
+\subsection{Proof of Theorem \ref{thm: padilla18_theorem_1}}
+We begin with a basic inequality
+\begin{equation*}
+\frac{1}{2} \norm{\widehat{\theta} - \theta^*}_n^2 \leq \frac{1}{n} \dotp{\epsilon}{\widehat{\theta} - \theta^{\star}} + \lambda_n \left(\norm{\nabla_G \theta^*}_1 - \norm{\nabla_G \widehat{\theta}}_1 \right)
+\end{equation*}
+
+We split up the \textcolor{red}{empirical process},
+\begin{equation*}
+\dotp{\epsilon}{\widehat{\theta} - \theta^{\star}} = \dotp{\epsilon}{\widehat{\theta} - \widehat{\theta}_I} + \dotp{\epsilon}{\widehat{\theta}_I - \theta_I^{\star}} + \dotp{\epsilon}{\theta_I^{\star} - \theta^{\star}}
+\end{equation*}
+Hereafter in the proof, we condition on the event $\Omega$. Lemma \ref{lem: quantization_error} gives us bounds on the first and third terms
+\begin{align*}
+\dotp{\epsilon}{\widehat{\theta} - \widehat{\theta}_I} & \leq 2 \norm{\epsilon}_{\infty} \cdot \norm{\nabla_{G_K} \widehat{\theta} }_1 \\
+\dotp{\epsilon}{\theta^{\star} - \theta_I^{\star}} & \leq 2 \norm{\epsilon}_{\infty} \cdot \norm{\nabla_{G_K} \theta^{\star} }_1
+\end{align*}
 

### 2019-02-15 notes/weekly_notes/2.14.19.NotesForWeek.tex
added updated version of this week's notes and notes on references

@@
 
-Consider $\Xbf = (X_1, \ldots, X_n) \sim \Pbb$ and $\Ybf = (Y_1, \ldots, Y_m) \sim \Qbb$ independently generated random variables.
-We will assume $\Pbb$ and $\Qbb$ are supported on $\D \subset \Reals^d$ and are absolutely continuous with density functions $f$ and $g$, respectively.
+Consider distributions $\Pbb$ and $\Qbb$ supported on $\D \subset \Reals^d$ which are absolutely continuous with density functions $f$ and $g$, respectively.
+For fixed $t \geq 0$, Let $\Zbf = (z_1, \ldots,z_t)$, where for $i = 1,\ldots,t$, $z_i \sim \frac{\Pbb + \Qbb}{2}$ are independent.
+Given $\Zbf$, for $i = 1,...,t$ let
+\begin{equation*}
+\ell_i = 
+\begin{cases}
+1~ \text{with probability $\frac{f(z_i)}{f(z_i) + g(z_i)}$} \\
+-1~ \text{with probability $\frac{g(z_i)}{f(z_i) + g(z_i)}$}
+\end{cases}
+\end{equation*} 
+be conditional independent labels, and write
+\begin{equation*}
+1_X = 
+\begin{cases}
+1,~ l_i = 1\\
+0,~ \text{otherwise}
+\end{cases}
+1_Y = 
+\begin{cases}
+1,~ l_i = -1 \\
+0,~ \text{otherwise.}
+\end{cases}
+\end{equation*}
+We will write $\Xbf = \set{x_1, \ldots,x_{N_1}} := \set{z_i: \ell_i = 1}$ and similarly $\Ybf = \set{y_1, \ldots,y_{N_2}} := \set{y_i: \ell_i = -1}$, where $N_1$ and $N_2$ are random.
+
 Our statistical goal is hypothesis testing: that is, we wish to construct a test function $\phi$ which differentiates between
@@
 \begin{equation*}
-R_{\epsilon}^{m,n}(\phi; \Hclass) = \sup_{f \in \Hclass} \Ebb_{f,f}^{(m,n)}(\phi) + \sup_{ \substack{f,g \in \Hclass \\ \delta(f,g) \geq \epsilon } } \Ebb_{f,g}^{m,n}(1 - \phi)
+R_{\epsilon}^{(t)}(\phi; \Hclass) = \sup_{f \in \Hclass} \Ebb_{f,f}^{(t)}(\phi) + \sup_{ \substack{f,g \in \Hclass \\ \delta(f,g) \geq \epsilon } } \Ebb_{f,g}^{(t)}(1 - \phi)
 \end{equation*}
@@
 
-Let $\Zbf = (X_1, \ldots, X_n, Y_1, \ldots, Y_m)$, and let $t = n + m$.
 As in \citep{padilla2018}, define the \emph{$K$-NN graph} $G_K = (V,E_K)$ to have vertex set $V = \set{1, \ldots, t}$ and edge set $E_K$ which contains the pair $(i,j)$ if and only if $x_i$ is among the $K$-nearest neighbors (with respect to Euclidean distance) of $x_j$, or vice versa.
@@
 \label{eqn: knn_tv_test_statistic}
-T_{TV} = \sup_{\substack{\theta \in \Reals^{t}: \\ \norm{D_{k} \theta}_1 \leq C_{n,k}} } \left(\sum_{i = 1}^{n} \theta_i - \sum_{j = n+1}^{t} \theta_j\right)
+T_{TV} = \sup_{\substack{\theta \in \Reals^{t}: \\ \norm{D_{k} \theta}_1 \leq C_{n,k}} } \left(\sum_{i: (1_X)_i = 1} \theta_i - \sum_{j:(1_Y)_j = 1} \theta_j\right)
 \end{equation}
@@
 	\begin{equation*}
-	R_{\epsilon}^{(m,m)}(\Hclass_{lip}(L)) \leq c_1/a^2
+	R_{\epsilon}^{(t)}(\Hclass_{lip}(L)) \leq c_1/a^2
 	\end{equation*}
@@
 	\begin{equation*}
-	\sum_{i = 1}^{m} \theta_i - \sum_{j = m + 1}^{2m} \theta_j = \dotp{\theta}{1_X - 1_Y}
+	\left(\sum_{i: (1_X)_i = 1} \theta_i - \sum_{j:(1_Y)_j = 1} \theta_j\right) = \dotp{\theta}{1_X - 1_Y}
 	\end{equation*}
-	where
-	\begin{equation*}
-	1_X = 
-	\begin{cases}
-	1,~ i = 1, \ldots,m \\
-	0,~ \text{otherwise}
-	\end{cases}
-	1_Y = 
-	\begin{cases}
-	1,~ i = j + 1, \ldots,2m \\
-	0,~ \text{otherwise}
-	\end{cases}
-	\end{equation*}
-	Let
+	and let
 	\begin{equation*}
@@
 
-	We bound the first term, and begin by employing a basic inequality argument. Let
+	 Let
 	\begin{align*}
@@
 	\end{align*}
+	We bound the first term, dividing into two cases.
+	\paragraph{Case 1:} First, suppose $\abs{\dotp{\widehat{\theta} - \theta^{\star}}{1_X - 1_Y}} \geq \norm{\theta - \widetilde{\theta}}_2^2$.
+
+	We begin by employing a basic inequality argument.
 	Since $\widetilde{\theta}$ is feasible and $\widehat{\theta}$ optimal for \eqref{eqn: knn_tv_test_statistic}, we have
@@
 	where $(i)$ follows from \eqref{eqn: basic_inequality_1}.
+
+	$w$ are mean-zero, bounded random variables, and are therefore subgaussian with parameter $2$. Then, from \citep{padilla2018}, we have
+	\begin{lemma}[Approximate restatement of \citep{padilla2018}]
+		\begin{equation}
+		\label{eqn: empirical_process_lower_bound}
+		\abs{\dotp{\widehat{\theta} - \theta^{\star}}{w}} \lesssim \sqrt{K \log n} \left[ \norm{D_K \widetilde{\theta}}_1 + \norm{D_K \widehat{\theta}} \right] + \sqrt{K} \cdot \left(\norm{\widehat{\theta} - \widetilde{\theta}}_2 \right)
+		\end{equation}
+		with probability at least 
+		\begin{equation*}
+		\Omega \left(1 - \frac{n}{K} \exp\left(-K\right) - n \exp(-K)  \right)
+		\end{equation*}
+	\end{lemma}
+	Condition on the event of \eqref{eqn: empirical_process_lower_bound} holding. As a result, we have
+	\begin{align*}
+	\abs{\dotp{\widehat{\theta} - \theta^{\star}}{1_X - 1_Y}} & \lesssim  \abs{\dotp{\widetilde{\theta} - \theta^{\star}}{\theta^{\star}} } + \sqrt{K \log n} \left[ \norm{D_K \widetilde{\theta}}_1 + \norm{D_K \widehat{\theta}} \right] + \sqrt{K} \cdot \left(\norm{\widehat{\theta} - \widetilde{\theta}}_2 \right) \\
+	& \leq \abs{\dotp{\widetilde{\theta} - \theta^{\star}}{\theta^{\star}} } + \sqrt{K \log n} \left[ \norm{D_K \widetilde{\theta}}_1 + \norm{D_K \widehat{\theta}} \right] + \sqrt{K} \cdot \left(\sqrt{\abs{\dotp{\widehat{\theta} - \theta^{\star}}{1_X - 1_Y}}} \right)
+	\end{align*}
+	which, along with the inequality $ab - b^2/4 \leq a^2$, implies
+	\begin{equation*}
+	\abs{\dotp{\widehat{\theta} - \theta^{\star}}{1_X - 1_Y}} \leq \abs{\dotp{\widetilde{\theta} - \theta^{\star}}{\theta^{\star}} } + \sqrt{K \log n} \left[ \norm{D_K \widetilde{\theta}}_1 + \norm{D_K \widehat{\theta}} \right] + K
+	\end{equation*}
+
+	\textcolor{red}{Continue from here.}
+
+	\paragraph{Case 2:} Otherwise, we may assume $\abs{\dotp{\widehat{\theta} - \theta^{\star}}{1_X - 1_Y}} \leq \norm{\theta - \widetilde{\theta}}_2^2$.
+
+	To upper bound $\norm{\theta - \widetilde{\theta}}_2^2$, we use the relation
+	\begin{equation*}
+	\dotp{\widehat{\theta}}{1_X - 1_Y} \leq  \dotp{\widetilde{\theta}}{1_X - 1_Y}
+	\end{equation*}
+	and therefore
+	\begin{equation*}
+	\norm{\widehat{\theta} - (1_X - 1_Y)}_2^2 \leq \norm{\widetilde{\theta} - (1_X - 1_Y)}_2^2 + \left(\norm{\widehat{\theta}}_2^2 - \norm{\widetilde{\theta}}_2^2\right)
+	\end{equation*}
+
+	\textcolor{red}{Continue from here using standard basic inequality + Poincare bound on the second term.}
+
 \end{proof}

### 2019-02-22 notes/weekly_notes/2.21.19.NotesForWeek.tex
added notes for the week

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbb{I}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Matrices
\newcommand{\Xbf}{\mathbf{X}}
\newcommand{\Ybf}{\mathbf{Y}}
\newcommand{\Zbf}{\mathbf{Z}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 2/16/19 - 2/22/19}
\author{Alden Green}
\date{\today}
\maketitle

Consider distributions $\Pbb$ and $\Qbb$ supported on $\D \subset \Reals^d$ which are absolutely continuous with density functions $f$ and $g$, respectively. For fixed $t \geq 0$, Let $\Zbf = (z_1, \ldots,z_t)$, where for $i = 1,\ldots,t$, $z_i \sim \frac{\Pbb + \Qbb}{2}$ are independent. Given $\Zbf$, for $i = 1,...,t$ let
\begin{equation*}
\ell_i = 
\begin{cases}
1~ \text{with probability $\frac{f(z_i)}{f(z_i) + g(z_i)}$} \\
-1~ \text{with probability $\frac{g(z_i)}{f(z_i) + g(z_i)}$}
\end{cases}
\end{equation*} 
be conditional independent labels, and write
\begin{equation*}
1_X = 
\begin{cases}
1,~ l_i = 1\\
0,~ \text{otherwise}
\end{cases}
1_Y = 
\begin{cases}
1,~ l_i = -1 \\
0,~ \text{otherwise.}
\end{cases}
\end{equation*}
We will write $\Xbf = \set{x_1, \ldots,x_{N_X}} := \set{z_i: \ell_i = 1}$ and similarly $\Ybf = \set{y_1, \ldots,y_{N_Y}} := \set{y_i: \ell_i = -1}$, where $N_X$ and $N_Y$ are of course random.

Our statistical goal is hypothesis testing: that is, we wish to construct a test function $\phi$ which differentiates between
\begin{equation*}
\mathbb{H}_0: f = g \text{ and } \mathbb{H}_1: f \neq g.
\end{equation*}

For a given function class $\Hclass$, some $\epsilon > 0$, and test function $\phi$ a Borel measurable function of the data with range $\set{0,1}$, we evaluate the quality of the test using \emph{worst-case risk}
\begin{equation*}
R_{\epsilon}^{(t)}(\phi; \Hclass) = \sup_{f \in \Hclass} \Ebb_{f,f}^{(t)}(\phi) + \sup_{ \substack{f,g \in \Hclass \\ \delta(f,g) \geq \epsilon } } \Ebb_{f,g}^{(t)}(1 - \phi)
\end{equation*} 
where 
\begin{equation*}
\delta^2(f,g) = \int_{\D} (f - g)^2 dx.
\end{equation*}

\section{Total variation test}

As in \citep{padilla2018}, define the \emph{$K$-NN graph} $G_K = (V,E_K)$ to have vertex set $V = \set{1, \ldots, t}$ and edge set $E_K$ which contains the pair $(i,j)$ if and only if $x_i$ is among the $K$-nearest neighbors (with respect to Euclidean distance) of $x_j$, or vice versa. Let $D_K$ denote the incidence matrix of $G_K$.

Define the \emph{kNN-total variation} test statistic to be
\begin{equation}
\label{eqn: knn_tv_test_statistic}
T_{TV} = \sup_{\substack{\theta \in \Reals^{t}: \\ \mathcal{T}(C_{n,k})} } \left(\frac{1}{N_X}\sum_{i: (1_X)_i = 1} \theta_i - \frac{1}{N_Y}\sum_{j:(1_Y)_j = 1} \theta_j\right)
\end{equation}
where $\mathcal{T}(C_{n,k}) = \set{\theta: \norm{D_{G_K}\theta}_1 \leq C_{n,k},
\norm{\theta}_2 \leq C'_{n,k}}$. 
Hereafter, take $\D = [0,1]^d$, and consider
\begin{equation*}
\Hclass_{lip}(L) = \set{f: [0,1]^d \to \Reals^{+}: \int_{\D}f = 1, f~ \text{$L$-piecewise Lipschitz, bounded above and below} }
\end{equation*}

\begin{definition}[Piecewise Lipschitz]
	\label{def: piecewise_lipschitz}
	A function $f$ is $L$-\emph{piecewise lipschitz} over $[0,1]^d$ if there exists a set $\Sset \subset [0,1]^d$ such that
	\begin{enumerate}[(a)]
		\item $\nu(\Sset) = 0$
		\item There exist $C_{\Sset}, \epsilon_0$ such that $\mu \Bigl( \bigl(\Sset_{\epsilon} \cup (\partial \D)_{\epsilon} \bigr) \cap [0,1]^d \Bigr) \leq C_{\Sset} \epsilon$ for all $0 < \epsilon \leq \epsilon_0$.
		\item For any $z, z'$ in the same connected component of $[0,1]^d \setminus \bigl(\Sset_{\epsilon} \cup (\partial \D)_{\epsilon}\bigr)$,
		\begin{equation*}
		\abs{g(z) - g(z')}_2 \leq L\norm{z - z'}_2
		\end{equation*}
	\end{enumerate}
\end{definition}

\begin{definition}[Bounded above and below]
	\label{def: bounded}
	A function $f: \D \to \Reals$ is \emph{bounded above and below} if there exists $p_{\min}, p_{\max}$ such that
	\begin{equation*}
	0 < p_{\min} < f(x) < p_{\max} < \infty \tag{$\forall x \in \D$}
	\end{equation*}
\end{definition}

\begin{conjecture}
	For $\tau = ???$ and $K \asymp \log^{1 + 2r}(n)$ for some $r \geq 0$, the test $\phi_{TV} = \set{T_{TV} \geq \tau}$ has worst-case risk
	\begin{equation*}
	R_{\epsilon}^{(t)}(\Hclass_{lip}(L)) \leq 1/2
	\end{equation*}
	whenever $\epsilon \geq c_2 \log^{\alpha}m m^{-1/d}$ where $\alpha = 3r + 5/2 + (2r + 1)/d$ and $c_1$ and $c_2$ are constants which depend only on $(d,L)$.
\end{conjecture}
\begin{proof}
	Write
	\begin{equation*}
	\left(\frac{1}{N_X}\sum_{i: (1_X)_i = 1} \theta_i - \frac{1}{N_Y}\sum_{j:(1_Y)_j = 1} \theta_j\right) = \dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
	\end{equation*}
	and let
	\begin{equation*}
	\widehat{\theta} \in \argmax_{\theta \in \Reals^{t}} \set{\dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}: \theta \in \mathcal{T}(C_{n,K})}
	\end{equation*}
	satisfy $T_{TV} = \dotp{\widehat{\theta}}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}$.

	\paragraph{Random denominators}

	We will first account for the dependence on random denominators $N_X$ and $N_Y$. For arbitrary $\theta \in \mathcal{T}(C_{n,K})$
	\begin{align*}
	\dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}} = \frac{2}{t}\dotp{\theta}{1_X - 1_Y} + 2\left(\frac{N_X - t/2}{(N_X)t}\right) \dotp{\theta}{1_X} + 2\left(\frac{t/2 - N_Y}{(N_Y)t}\right) \dotp{\theta}{1_Y}
	\end{align*}
	A simple application of Holder's inequality, and standard concentration results for binomial random variables, will show the latter two terms to be $O_{P}\left(\frac{C'_{n,k}}{t}\right)$.

	We turn to analyzing the first term. For $i = 1, \ldots, t$, introduce $\theta^{\star}$ defined by
	\begin{equation*}
	(\theta^{\star})_i = \left(\frac{f(z_i) - g(z_i)}{f(z_i) + g(z_i)} \right)	\end{equation*}
	Letting $w = (1_X - 1_Y) - \theta^{\star}$, note that $w$ is a vector of i.i.d bounded random variables with $\Ebb(w) = 0$.

	\paragraph{Type I error.}
	Under the case $f = g$, we have $\theta^{\star} = 0$, and therefore
	\begin{equation*}
	\dotp{\theta}{1_X - 1_Y} = \dotp{\theta}{w}
	\end{equation*}

	\textcolor{red}{Proceed using the empirical process bound of Lemma \ref{lem: empirical_process_TV_denoising}.}
\end{proof}
\section{Bounding the empirical process}

We collect here the bound on the empirical process from \cite{padilla2018}.
\begin{lemma}
	\label{lem: empirical_process_TV_denoising}
	Let $w$ be a vector of mean zero independent random variables with ${w}_{\infty} \leq 1$. Then, for any $\delta > 0$ such that $K \geq 3 \log(n/\delta)$
	\begin{equation*}
	\sup_{\theta \in \mathcal{T}(C_{n,K})} \dotp{\theta}{w} \leq 2 C_{n,K} + \sqrt{(1 + \frac{C(p_{\max}, \delta)}{K})K p_{\max}} \cdot \left( 2 \sqrt{2 \log(e/ \delta)} C'_{n,K} + 2 C(d) \sqrt{\log(en/\delta)}C_{n,K} \right)
	\end{equation*}
	with probability at least $1 - 4 \delta$.
\end{lemma}
\begin{proof}
	Set 
	\begin{equation*}
	\kappa = \lceil \frac{3\sqrt{d} p_{\max}^{1/d}t^{1/d}}{2 K^{1/d}} \rceil
	\end{equation*}
	and let 
	\begin{equation*}
	I = \set{\biggl(\frac{k - 1}{\kappa}, \frac{k}{\kappa}\biggr]: k \in [\kappa]^d},~ M_{k, \kappa} = \# \set{i \in [t]: z_i \in \biggl(\frac{k - 1}{\kappa}, \frac{k}{\kappa}\biggr]}
	\end{equation*}
	be a partition of $\D$ into cells, and the count in each cell, respectively. From \citep{padilla2018} we have that with probability at least
	\begin{equation*}
	1 - t \exp(-K/3) \geq 1 - \delta
	\end{equation*}
	the following statement holds:
	\begin{equation*}
	\sup_{\theta \in \mathcal{T}(C_{n,K})} \leq 2 \norm{w}_{\infty} \norm{D_{G_K} \theta}_1 + \max_{k} \sqrt{M_{k,\kappa}} \left( \norm{\Pi \widetilde{w}}_2 \norm{\theta}_2 + \norm{(D^{\dagger})^T \widetilde{w}}_{\infty} \norm{D_{G_K} \theta}_1 \right)
	\end{equation*}
	where $\Pi$ is the projection onto the span of $1_{\kappa^d}$, $D$ is the incidence matrix of the grid graph $\kappa^d$, and $\widetilde{w}$ is defined by
	\begin{equation*}
	(\widetilde{w})_k = \left[\max_{k \in [\kappa]^d} M_{k,\kappa}\right]^{-1/2} \sum_{l: z_l \in (\frac{k - 1}{\kappa}, \frac{k}{\kappa}]} w_{l}
	\end{equation*}
	\textcolor{red}{Take care of random denominator again.}

	\citep{hutter2016} derive the following bounds, which hold with probability at least $1 - 2\delta$:
	\begin{equation*}
	\norm{(D^{\dagger})^T w}_{\infty} \leq 2 C(d) \sqrt{\log(en/\delta)}, ~~ \norm{\Pi w}_2 \leq 2 \sqrt{2 \log(e / \delta)}
	\end{equation*}

	Then, a simple concentration inequality for binomial random variables, along with a union bound, gives
	\begin{equation*}
	\max_{k \in [\kappa]^d} \sqrt{M_{k, \kappa}} \leq \sqrt{(1 + \frac{C(p_{\max}, \delta)}{K})\frac{tp_{\max}}{\kappa^d}}
	\end{equation*}
	with probability at least $1 - \delta$. The statement follows from our choice of $\kappa$.
\end{proof}

\section{Laplacian smooth test--Attempt 1}
Let
\begin{equation*}
\widehat{\theta}_{LS} = \sup_{\theta \in \mathcal{S}(C_{n,K}, C'_{n,K})} \dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
\end{equation*}
where $\mathcal{S}(C_{n,K}, C'_{n,K}) = \set{\theta \in \Reals^n: \norm{D_{G_K\theta}}_2 \leq C_{n,K}, \norm{\theta}_2 \leq C'_{n,K}}$.

\paragraph{Random Denominator.}
\begin{lemma}
	Let $\norm{\theta}_2 \leq c$ and $\delta > 0$. Then,
	\begin{equation*}
	\abs{ \dotp{\theta}{1_X} \left(\frac{1}{N_X} - \frac{2}{t}\right) } \vee \abs{ \dotp{\theta}{1_Y} \left(\frac{1}{N_Y} - \frac{2}{t}\right) } \leq \frac{c \sqrt{\log(2 / \delta)}}{t (1 - \log(2/\delta)/\sqrt{t})}
	\end{equation*}
	with probability at least $1 - \delta$.
\end{lemma}
\begin{proof}
	By the Cauchy-Schwarz inequality,
	\begin{equation*}
	\abs{\dotp{\theta}{1_X}} \leq \norm{\theta}_2 \norm{1_X}_2 \leq c \sqrt{N_X}.
	\end{equation*}
	Rearranging $1/N_X - 2/t$, we obtain
	\begin{equation}
	\label{eqn: random_denominator_1}
	\abs{ \dotp{\theta}{1_X} \left(\frac{1}{N_X} - \frac{2}{t}\right) } \leq 2 \frac{c \sqrt{N_X} \abs{N_X - \frac{2}{t}}}{tN_X} = \frac{c \abs{N_X - \frac{2}{t}}}{t\sqrt{N_X}} 
	\end{equation}
	$N_X \sim Bin(t,1/2)$, and so application of Hoeffding's inequality gives
	\begin{equation}
	\label{eqn: hoeffding_bound_nx}
	\abs{N_X - \frac{t}{2}} \leq \sqrt{t} \sqrt{\log(2/\delta)}
	\end{equation}
	with probability at least $1 - \delta$. Plugging this in to \eqref{eqn: random_denominator_1} yields the desired bound with respect to $1_X,N_X$. However, if \eqref{eqn: hoeffding_bound_nx} holds for $N_X$ it holds for $N_Y$ as well. All other steps hold for $1_Y$, and therefore the desired bound holds with respect to $1_Y, N_Y$ as well.
\end{proof}

\section{Laplacian smooth test: Attempt 2}
Let
\begin{equation*}
T_{LS} = \sup_{\theta \in \mathcal{S}(C_{n,K}, C'_{n,K})} \dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
\end{equation*}
where $\mathcal{S}(C_{n,K}, C'_{n,K}) = \set{\theta \in \Reals^n: \norm{D_{G_K\theta}}_2 \leq C_{n,K}}$. We can find a closed-form solution to this problem,
\begin{align*}
T_{LS} & = C_{n,k} a^T \Linv a
\end{align*}
where $a = (\frac{1_X}{N_X} - \frac{1_Y}{N_Y})$.

\paragraph{Type I error.}

To begin, we write $w = (1_X - 1_Y)\frac{2}{t}$, and rewrite
\begin{equation*}
T_{LS} = C_{n,k} w^T \Linv w + C_{n,k} (w + \ell)^T \Linv (\ell - w)
\end{equation*}

We turn our attention to the second term, which we wish to show contributes negligibly to the overall sum. We have
\begin{align*}
(w + \ell)^T \Linv (\ell - w) & = (D(w + \ell))^T (D(\ell - w)) \\
& \leq \norm{D(w + \ell)}_2 \norm{D(w - \ell)}_2 \\
& \leq K \norm{w + \ell}_2 \norm{w - \ell}_2
\end{align*}
Then, based on Lemma \textcolor{red}{make a `random denominators` Lemma}, with probability at least $1 - \delta$,
\begin{equation*}
\norm{w + \ell}_2 \leq \frac{1}{\sqrt{t}}, ~~ \norm{w - \ell}_2 \leq 2 \frac{\log(2/\delta)}{t \left(1 - \frac{\log(2/\delta)}{\sqrt{t}}\right)}
\end{equation*}
and as a result
\begin{equation*}
(w + \ell)^T \Linv (\ell - w) \leq K \frac{\log(2/\delta)}{t^{3/2} \left(1 - \frac{\log(2/\delta)}{\sqrt{t}}\right)}
\end{equation*}

Now, we have that the entries of $w$ are i.i.d random variables with mean $0$ and absolute value of $2/t$.

\section{Laplacian Smooth Test-Attempt 3}
Define the \emph{$r$-graph} $G_r = (V,E_r)$ to have vertex set $V = \set{1,\ldots,t}$ and edge set $E_r$ which contains the pair $(i,j)$ if and only if $\norm{z_i - z_j}_2 \leq r$. Let $D_{G_r}$ denote the incidence matrix of $G_r$.

Define the \emph{$r$-Laplacian smooth} test statistic to be
\begin{equation*}
T_{LS} = \sup_{\theta: \norm{D_{G_r}\theta}_2 \leq C_{n,r}} \dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
\end{equation*}

We would like to relate the graph $G_r$ to a graph with a more easily accessible spectrum. For $\kappa = t^{1/d}$, consider the \emph{grid graph}
\begin{equation*}
G_{grid} = (V_{grid},E_{grid}),~~ V_{grid} = \set{\frac{k}{\kappa}: k \in [\kappa]^d},~~ E_{grid} = \set{(k,k'): k, k' \in V_{grid}, \norm{k - k'}_1 = \frac{1}{\kappa^d}}
\end{equation*}
with associated incidence matrix $D_{grid}$.

\begin{lemma}[Spectral similarity of $r$-graph to grid]
	\label{lem: spectral_similarity_rgraph_grid}
	Fix $r \geq 2 \left(\frac{\log t}{t}\right)^{1/d} + (\frac{1}{t})^{1/d}$, and Let $\ell(t) = \sqrt{d} r t^{1/d} + 2\sqrt{d}(\log t)^{1/d}$. For any $\theta \in \Reals^t$, the following relations hold:
	\begin{equation}
	\label{eqn: spectral_similarity_rgraph_grid}
	\frac{\norm{D_{G_r}\theta}_2}{\ell(t)} \leq \norm{D_{grid}\theta}_2 \leq \norm{D_{G_r}\theta}_2
	\end{equation}
	with probability at least $1 - n^{-\alpha}$ where $\alpha = c_1 (\log n)^{1/2}$ for some constant $c_1 > 0$.
\end{lemma}
\begin{proof}
	We begin by mapping the data $\Zbf$ to the grid points $[\kappa]^d$ in such a way that as little mass as possible is disturbed:
	\begin{lemma}
		\label{lem: mass_transport_mapping}
		There exists a bijective mapping $T: \Zbf \to [\kappa]^d$ for $\kappa = t^{1/d}$ such that
		\begin{equation*}
		\max_{i} \norm{T(z_i) - z_i}_2 \leq C\left(\frac{\log t}{t}\right)^{1/d}
		\end{equation*}
		with probability at least $1 - n^{-\alpha}$ where $\alpha = c_1 (\log n)^{1/2}$ for some constant $c_1 > 0$.
	\end{lemma}
	Hereafter, we assume there exists $T$ such that Lemma \ref{lem: mass_transport_mapping} holds.

	We first prove the second bound in \eqref{eqn: spectral_similarity_rgraph_grid}.  Consider grid points $k ~ k'$ connected in the grid graph. Then, there exist $z_i$ and $z_j$ such that $T(z_i) = k$ and $T(z_j) = k'$. By the triangle inequality,
	\begin{align*}
	\norm{z_i - z_j}_2 & \leq \norm{T(z_i) - z_i}_2 + \norm{T(z_i) - T(z_j)}_2 + \norm{T(z_j) - z_j}_2 \\
	& \leq 2C\left(\frac{\log t}{t}\right)^{1/d} + \frac{1}{t^{1/d}}
	\end{align*}
	and so by our choice of $r$, $i \sim j$ in $G_r$.

	Now, we turn to the first bound. Assume $i \sim j$ in the graph $G_r$. By a similar set of steps to the above, we have
	\begin{equation*}
	\norm{T(z_i) - T(z_j)}_2 \leq 2C\left(\frac{\log t}{t}\right)^{1/d} + r
	\end{equation*}
	As a result, using the simple relation $\norm{x}_1 \leq \sqrt{d} \norm{x}_2$ for any $x \in \Rd$, we have
	\begin{equation*}
	\norm{T(z_i) - T(z_j)}_1 \leq \sqrt{d}(2C\left(\frac{\log t}{t}\right)^{1/d} + r)
	\end{equation*}
	Since each edge in the grid graph is of length $n^{1/d}$, it is easy to see that there exists a path between $T(z_i)$ and $T(z_j)$ in $G_{grid}$, $P(T(Z_i) \to T(Z_j))$ with no more than
	\begin{equation*}
		\frac{\sqrt{d}(2C\left(\frac{\log t}{t}\right)^{1/d} + r)}{t^{1/d}}
	\end{equation*}
	edges. The bound follows by Lemma
	\begin{lemma}[Graph ordering]
	\label{lem: graph_ordering}
	Fix $m \geq 0$. For vertices $V = \set{1, \ldots,m}$, we have
	\begin{enumerate}
		\item $\frac{1}{m - 1}P(1 \to m) \succeq G_{1,m}$
		\item If $A \succeq B$ and $C \succeq D$, then $A + B \succeq C + D$. 
	\end{enumerate}
	\end{lemma}
\end{proof}

Decompose $\frac{1_X}{N_X} - \frac{1_Y}{N_Y} := \theta^{\star} + w$, where
\begin{equation*}
(\theta^{\star})_i := \frac{f(x) - g(x)}{f(x) + g(x)}
\end{equation*}

The upper bound in Lemma \ref{lem: spectral_similarity_rgraph_grid} allows us the following upper bound on the empirical process
\begin{equation*}
\sup_{\theta: \norm{D_{r}\theta}_2 \leq C_{n,r}} \dotp{\theta}{w} \leq  \sup_{\theta: \norm{D_{grid}\theta}_2 \leq C_{n,r}} \dotp{\theta}{w} = C_{n,r} w^T{\Linv_{grid}}w 
\end{equation*}
whereas the lower bound helps us with the approximation error term,
\begin{equation*}
\sup_{\widetilde{\theta}: \norm{D_{r}\theta}_2 \leq C_{n,r}} \dotp{\widetilde{\theta}}{\theta^{\star}} \geq \sup_{\theta: \norm{D_{grid}\theta}_2 \leq C_{n,r}/ \ell(n,r)} \dotp{\theta}{\theta^{\star}} \geq \frac{C_{n,r}}{\ell(n,r)} \theta^{\star}\Linv_{grid} \theta^{\star}
\end{equation*}

\clearpage

\bibliography{../../graph_testing_bibliography}
\bibliographystyle{plain}

\end{document}

### 2019-02-22 notes/weekly_notes/2.21.19.NotesForWeek.tex
made one or two changes

@@
 	\label{lem: spectral_similarity_rgraph_grid}
-	Fix $r \geq 2 \left(\frac{\log t}{t}\right)^{1/d} + (\frac{1}{t})^{1/d}$, and Let $\ell(t) = \sqrt{d} r t^{1/d} + 2\sqrt{d}(\log t)^{1/d}$.
+	Fix $r \geq 2 \left(\frac{\log t}{t}\right)^{1/d} + (\frac{1}{t})^{1/d}$, and let $\ell(t) = \sqrt{\sqrt{d} r t^{1/d} + 2\sqrt{d}(\log t)^{1/d}}$.
 For any $\theta \in \Reals^t$, the following relations hold:
@@
 	\end{equation*}
-	edges. The bound follows by Lemma
+	edges. The bound follows by Lemma \ref{lem: graph_ordering}.
 	\begin{lemma}[Graph ordering]

### 2019-02-23 notes/weekly_notes/2.29.19.NotesForWeek.tex
started this week's notes

--- 2019-02-22 notes/weekly_notes/2.21.19.NotesForWeek.tex
@@
 \begin{document}
-\title{Notes for Week 2/16/19 - 2/22/19}
+\title{Notes for Week 2/23/19 - 2/29/19}
 \author{Alden Green}
@@
 Consider distributions $\Pbb$ and $\Qbb$ supported on $\D \subset \Reals^d$ which are absolutely continuous with density functions $f$ and $g$, respectively.
-For fixed $t \geq 0$, Let $\Zbf = (z_1, \ldots,z_t)$, where for $i = 1,\ldots,t$, $z_i \sim \frac{\Pbb + \Qbb}{2}$ are independent.
+For fixed $n \geq 0$, Let $\Zbf = (z_1, \ldots,z_n)$, where for $i = 1,\ldots,t$, $z_i \sim \frac{\Pbb + \Qbb}{2}$ are independent.
 Given $\Zbf$, for $i = 1,...,t$ let
@@
 \end{equation*}
-We will write $\Xbf = \set{x_1, \ldots,x_{N_X}} := \set{z_i: \ell_i = 1}$ and similarly $\Ybf = \set{y_1, \ldots,y_{N_Y}} := \set{y_i: \ell_i = -1}$, where $N_X$ and $N_Y$ are of course random.
+We will write $\Xbf = \set{x_1, \ldots,x_{N_X}} := \set{z_i: \ell_i = 1}$ and similarly $\Ybf = \set{y_1, \ldots,y_{N_Y}} := \set{y_i: \ell_i = -1}$, where $N_X$ and $N_Y$ are of course random but $N_X + N_Y = n$.
 
@@
 
-\section{Total variation test}
+\section{Laplacian smooth test statistic}
 
-As in \citep{padilla2018}, define the \emph{$K$-NN graph} $G_K = (V,E_K)$ to have vertex set $V = \set{1, \ldots, t}$ and edge set $E_K$ which contains the pair $(i,j)$ if and only if $x_i$ is among the $K$-nearest neighbors (with respect to Euclidean distance) of $x_j$, or vice versa.
-Let $D_K$ denote the incidence matrix of $G_K$.
-
-Define the \emph{kNN-total variation} test statistic to be
-\begin{equation}
-\label{eqn: knn_tv_test_statistic}
-T_{TV} = \sup_{\substack{\theta \in \Reals^{t}: \\ \mathcal{T}(C_{n,k})} } \left(\frac{1}{N_X}\sum_{i: (1_X)_i = 1} \theta_i - \frac{1}{N_Y}\sum_{j:(1_Y)_j = 1} \theta_j\right)
-\end{equation}
-where $\mathcal{T}(C_{n,k}) = \set{\theta: \norm{D_{G_K}\theta}_1 \leq C_{n,k},
-\norm{\theta}_2 \leq C'_{n,k}}$. 
-Hereafter, take $\D = [0,1]^d$, and consider
-\begin{equation*}
-\Hclass_{lip}(L) = \set{f: [0,1]^d \to \Reals^{+}: \int_{\D}f = 1, f~ \text{$L$-piecewise Lipschitz, bounded above and below} }
-\end{equation*}
-
-\begin{definition}[Piecewise Lipschitz]
-	\label{def: piecewise_lipschitz}
-	A function $f$ is $L$-\emph{piecewise lipschitz} over $[0,1]^d$ if there exists a set $\Sset \subset [0,1]^d$ such that
-	\begin{enumerate}[(a)]
-		\item $\nu(\Sset) = 0$
-		\item There exist $C_{\Sset}, \epsilon_0$ such that $\mu \Bigl( \bigl(\Sset_{\epsilon} \cup (\partial \D)_{\epsilon} \bigr) \cap [0,1]^d \Bigr) \leq C_{\Sset} \epsilon$ for all $0 < \epsilon \leq \epsilon_0$.
-		\item For any $z, z'$ in the same connected component of $[0,1]^d \setminus \bigl(\Sset_{\epsilon} \cup (\partial \D)_{\epsilon}\bigr)$,
-		\begin{equation*}
-		\abs{g(z) - g(z')}_2 \leq L\norm{z - z'}_2
-		\end{equation*}
-	\end{enumerate}
-\end{definition}
-
-\begin{definition}[Bounded above and below]
-	\label{def: bounded}
-	A function $f: \D \to \Reals$ is \emph{bounded above and below} if there exists $p_{\min}, p_{\max}$ such that
-	\begin{equation*}
-	0 < p_{\min} < f(x) < p_{\max} < \infty \tag{$\forall x \in \D$}
-	\end{equation*}
-\end{definition}
-
-\begin{conjecture}
-	For $\tau = ???$ and $K \asymp \log^{1 + 2r}(n)$ for some $r \geq 0$, the test $\phi_{TV} = \set{T_{TV} \geq \tau}$ has worst-case risk
-	\begin{equation*}
-	R_{\epsilon}^{(t)}(\Hclass_{lip}(L)) \leq 1/2
-	\end{equation*}
-	whenever $\epsilon \geq c_2 \log^{\alpha}m m^{-1/d}$ where $\alpha = 3r + 5/2 + (2r + 1)/d$ and $c_1$ and $c_2$ are constants which depend only on $(d,L)$.
-\end{conjecture}
-\begin{proof}
-	Write
-	\begin{equation*}
-	\left(\frac{1}{N_X}\sum_{i: (1_X)_i = 1} \theta_i - \frac{1}{N_Y}\sum_{j:(1_Y)_j = 1} \theta_j\right) = \dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
-	\end{equation*}
-	and let
-	\begin{equation*}
-	\widehat{\theta} \in \argmax_{\theta \in \Reals^{t}} \set{\dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}: \theta \in \mathcal{T}(C_{n,K})}
-	\end{equation*}
-	satisfy $T_{TV} = \dotp{\widehat{\theta}}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}$.
-
-	\paragraph{Random denominators}
-
-	We will first account for the dependence on random denominators $N_X$ and $N_Y$. For arbitrary $\theta \in \mathcal{T}(C_{n,K})$
-	\begin{align*}
-	\dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}} = \frac{2}{t}\dotp{\theta}{1_X - 1_Y} + 2\left(\frac{N_X - t/2}{(N_X)t}\right) \dotp{\theta}{1_X} + 2\left(\frac{t/2 - N_Y}{(N_Y)t}\right) \dotp{\theta}{1_Y}
-	\end{align*}
-	A simple application of Holder's inequality, and standard concentration results for binomial random variables, will show the latter two terms to be $O_{P}\left(\frac{C'_{n,k}}{t}\right)$.
-
-	We turn to analyzing the first term. For $i = 1, \ldots, t$, introduce $\theta^{\star}$ defined by
-	\begin{equation*}
-	(\theta^{\star})_i = \left(\frac{f(z_i) - g(z_i)}{f(z_i) + g(z_i)} \right)	\end{equation*}
-	Letting $w = (1_X - 1_Y) - \theta^{\star}$, note that $w$ is a vector of i.i.d bounded random variables with $\Ebb(w) = 0$.
-
-	\paragraph{Type I error.}
-	Under the case $f = g$, we have $\theta^{\star} = 0$, and therefore
-	\begin{equation*}
-	\dotp{\theta}{1_X - 1_Y} = \dotp{\theta}{w}
-	\end{equation*}
-
-	\textcolor{red}{Proceed using the empirical process bound of Lemma \ref{lem: empirical_process_TV_denoising}.}
-\end{proof}
-\section{Bounding the empirical process}
-
-We collect here the bound on the empirical process from \cite{padilla2018}.
-\begin{lemma}
-	\label{lem: empirical_process_TV_denoising}
-	Let $w$ be a vector of mean zero independent random variables with ${w}_{\infty} \leq 1$. Then, for any $\delta > 0$ such that $K \geq 3 \log(n/\delta)$
-	\begin{equation*}
-	\sup_{\theta \in \mathcal{T}(C_{n,K})} \dotp{\theta}{w} \leq 2 C_{n,K} + \sqrt{(1 + \frac{C(p_{\max}, \delta)}{K})K p_{\max}} \cdot \left( 2 \sqrt{2 \log(e/ \delta)} C'_{n,K} + 2 C(d) \sqrt{\log(en/\delta)}C_{n,K} \right)
-	\end{equation*}
-	with probability at least $1 - 4 \delta$.
-\end{lemma}
-\begin{proof}
-	Set 
-	\begin{equation*}
-	\kappa = \lceil \frac{3\sqrt{d} p_{\max}^{1/d}t^{1/d}}{2 K^{1/d}} \rceil
-	\end{equation*}
-	and let 
-	\begin{equation*}
-	I = \set{\biggl(\frac{k - 1}{\kappa}, \frac{k}{\kappa}\biggr]: k \in [\kappa]^d},~ M_{k, \kappa} = \# \set{i \in [t]: z_i \in \biggl(\frac{k - 1}{\kappa}, \frac{k}{\kappa}\biggr]}
-	\end{equation*}
-	be a partition of $\D$ into cells, and the count in each cell, respectively. From \citep{padilla2018} we have that with probability at least
-	\begin{equation*}
-	1 - t \exp(-K/3) \geq 1 - \delta
-	\end{equation*}
-	the following statement holds:
-	\begin{equation*}
-	\sup_{\theta \in \mathcal{T}(C_{n,K})} \leq 2 \norm{w}_{\infty} \norm{D_{G_K} \theta}_1 + \max_{k} \sqrt{M_{k,\kappa}} \left( \norm{\Pi \widetilde{w}}_2 \norm{\theta}_2 + \norm{(D^{\dagger})^T \widetilde{w}}_{\infty} \norm{D_{G_K} \theta}_1 \right)
-	\end{equation*}
-	where $\Pi$ is the projection onto the span of $1_{\kappa^d}$, $D$ is the incidence matrix of the grid graph $\kappa^d$, and $\widetilde{w}$ is defined by
-	\begin{equation*}
-	(\widetilde{w})_k = \left[\max_{k \in [\kappa]^d} M_{k,\kappa}\right]^{-1/2} \sum_{l: z_l \in (\frac{k - 1}{\kappa}, \frac{k}{\kappa}]} w_{l}
-	\end{equation*}
-	\textcolor{red}{Take care of random denominator again.}
-
-	\citep{hutter2016} derive the following bounds, which hold with probability at least $1 - 2\delta$:
-	\begin{equation*}
-	\norm{(D^{\dagger})^T w}_{\infty} \leq 2 C(d) \sqrt{\log(en/\delta)}, ~~ \norm{\Pi w}_2 \leq 2 \sqrt{2 \log(e / \delta)}
-	\end{equation*}
-
-	Then, a simple concentration inequality for binomial random variables, along with a union bound, gives
-	\begin{equation*}
-	\max_{k \in [\kappa]^d} \sqrt{M_{k, \kappa}} \leq \sqrt{(1 + \frac{C(p_{\max}, \delta)}{K})\frac{tp_{\max}}{\kappa^d}}
-	\end{equation*}
-	with probability at least $1 - \delta$. The statement follows from our choice of $\kappa$.
-\end{proof}
-
-\section{Laplacian smooth test--Attempt 1}
-Let
-\begin{equation*}
-\widehat{\theta}_{LS} = \sup_{\theta \in \mathcal{S}(C_{n,K}, C'_{n,K})} \dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
-\end{equation*}
-where $\mathcal{S}(C_{n,K}, C'_{n,K}) = \set{\theta \in \Reals^n: \norm{D_{G_K\theta}}_2 \leq C_{n,K}, \norm{\theta}_2 \leq C'_{n,K}}$.
-
-\paragraph{Random Denominator.}
-\begin{lemma}
-	Let $\norm{\theta}_2 \leq c$ and $\delta > 0$. Then,
-	\begin{equation*}
-	\abs{ \dotp{\theta}{1_X} \left(\frac{1}{N_X} - \frac{2}{t}\right) } \vee \abs{ \dotp{\theta}{1_Y} \left(\frac{1}{N_Y} - \frac{2}{t}\right) } \leq \frac{c \sqrt{\log(2 / \delta)}}{t (1 - \log(2/\delta)/\sqrt{t})}
-	\end{equation*}
-	with probability at least $1 - \delta$.
-\end{lemma}
-\begin{proof}
-	By the Cauchy-Schwarz inequality,
-	\begin{equation*}
-	\abs{\dotp{\theta}{1_X}} \leq \norm{\theta}_2 \norm{1_X}_2 \leq c \sqrt{N_X}.
-	\end{equation*}
-	Rearranging $1/N_X - 2/t$, we obtain
-	\begin{equation}
-	\label{eqn: random_denominator_1}
-	\abs{ \dotp{\theta}{1_X} \left(\frac{1}{N_X} - \frac{2}{t}\right) } \leq 2 \frac{c \sqrt{N_X} \abs{N_X - \frac{2}{t}}}{tN_X} = \frac{c \abs{N_X - \frac{2}{t}}}{t\sqrt{N_X}} 
-	\end{equation}
-	$N_X \sim Bin(t,1/2)$, and so application of Hoeffding's inequality gives
-	\begin{equation}
-	\label{eqn: hoeffding_bound_nx}
-	\abs{N_X - \frac{t}{2}} \leq \sqrt{t} \sqrt{\log(2/\delta)}
-	\end{equation}
-	with probability at least $1 - \delta$.
-Plugging this in to \eqref{eqn: random_denominator_1} yields the desired bound with respect to $1_X,N_X$.
-However, if \eqref{eqn: hoeffding_bound_nx} holds for $N_X$ it holds for $N_Y$ as well.
-All other steps hold for $1_Y$, and therefore the desired bound holds with respect to $1_Y, N_Y$ as well.
-\end{proof}
-
-\section{Laplacian smooth test: Attempt 2}
-Let
-\begin{equation*}
-T_{LS} = \sup_{\theta \in \mathcal{S}(C_{n,K}, C'_{n,K})} \dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
-\end{equation*}
-where $\mathcal{S}(C_{n,K}, C'_{n,K}) = \set{\theta \in \Reals^n: \norm{D_{G_K\theta}}_2 \leq C_{n,K}}$. We can find a closed-form solution to this problem,
-\begin{align*}
-T_{LS} & = C_{n,k} a^T \Linv a
-\end{align*}
-where $a = (\frac{1_X}{N_X} - \frac{1_Y}{N_Y})$.
-
-\paragraph{Type I error.}
-
-To begin, we write $w = (1_X - 1_Y)\frac{2}{t}$, and rewrite
-\begin{equation*}
-T_{LS} = C_{n,k} w^T \Linv w + C_{n,k} (w + \ell)^T \Linv (\ell - w)
-\end{equation*}
-
-We turn our attention to the second term, which we wish to show contributes negligibly to the overall sum. We have
-\begin{align*}
-(w + \ell)^T \Linv (\ell - w) & = (D(w + \ell))^T (D(\ell - w)) \\
-& \leq \norm{D(w + \ell)}_2 \norm{D(w - \ell)}_2 \\
-& \leq K \norm{w + \ell}_2 \norm{w - \ell}_2
-\end{align*}
-Then, based on Lemma \textcolor{red}{make a `random denominators` Lemma}, with probability at least $1 - \delta$,
-\begin{equation*}
-\norm{w + \ell}_2 \leq \frac{1}{\sqrt{t}}, ~~ \norm{w - \ell}_2 \leq 2 \frac{\log(2/\delta)}{t \left(1 - \frac{\log(2/\delta)}{\sqrt{t}}\right)}
-\end{equation*}
-and as a result
-\begin{equation*}
-(w + \ell)^T \Linv (\ell - w) \leq K \frac{\log(2/\delta)}{t^{3/2} \left(1 - \frac{\log(2/\delta)}{\sqrt{t}}\right)}
-\end{equation*}
-
-Now, we have that the entries of $w$ are i.i.d random variables with mean $0$ and absolute value of $2/t$.
-
-\section{Laplacian Smooth Test-Attempt 3}
-Define the \emph{$r$-graph} $G_r = (V,E_r)$ to have vertex set $V = \set{1,\ldots,t}$ and edge set $E_r$ which contains the pair $(i,j)$ if and only if $\norm{z_i - z_j}_2 \leq r$.
+For $r \geq 0$, define the \emph{$r$-graph} $G_r = (V,E_r)$ to have vertex set $V = \set{1,\ldots,t}$ and edge set $E_r$ which contains the pair $(i,j)$ if and only if $\norm{z_i - z_j}_2 \leq r$.
 Let $D_{G_r}$ denote the incidence matrix of $G_r$.
 
-Define the \emph{$r$-Laplacian smooth} test statistic to be
+Define the \emph{$r$-Laplacian Smooth} test statistic to be
 \begin{equation*}
@@
 
-We would like to relate the graph $G_r$ to a graph with a more easily accessible spectrum. For $\kappa = t^{1/d}$, consider the \emph{grid graph}
+We would like to relate the graph $G_r$ to a graph with a more easily accessible spectrum. For $\kappa = n^{1/d}$, consider the \emph{grid graph}
 \begin{equation*}
@@
 	\label{lem: spectral_similarity_rgraph_grid}
-	Fix $r \geq 2 \left(\frac{\log t}{t}\right)^{1/d} + (\frac{1}{t})^{1/d}$, and let $\ell(t) = \sqrt{\sqrt{d} r t^{1/d} + 2\sqrt{d}(\log t)^{1/d}}$.
-For any $\theta \in \Reals^t$, the following relations hold:
+	Fix $r \geq 2 \left(\frac{\log t}{n}\right)^{1/d} + (\frac{1}{n})^{1/d}$, and let $\ell(r,n) = \biggl(\sqrt{d} r n^{1/d} + 2\sqrt{d}(\log n)^{1/d}\biggr)^{1/2}$.
+For any $\theta \in \Reals^n$, the following relations hold:
 	\begin{equation}
@@
 	edges. The bound follows by Lemma \ref{lem: graph_ordering}.
-	\begin{lemma}[Graph ordering]
-	\label{lem: graph_ordering}
-	Fix $m \geq 0$. For vertices $V = \set{1, \ldots,m}$, we have
-	\begin{enumerate}
-		\item $\frac{1}{m - 1}P(1 \to m) \succeq G_{1,m}$
-		\item If $A \succeq B$ and $C \succeq D$, then $A + B \succeq C + D$. 
-	\end{enumerate}
+\end{proof}
+
+\section{Poincare inequalities}
+
+Let $G$ and $\widetilde{G}$ be undirected, unweighted graphs over vertex set $V$, with edge sets $E_G$ and $E_{\widetilde{G}}$, respectively.
+Let $\widetilde{\mathcal{P}}$ be the space of all paths over $E_{\widetilde{G}}$; that is, $\mathcal{P}$ consists of $\widetilde{P} \in \widetilde{\mathcal{P}}$
+\begin{equation*}
+P = (\widetilde{e}_1, \ldots, \widetilde{e}_m) \tag{$\widetilde{e_i} \in E_{\widetilde{G}}$}
+\end{equation*}
+for some integer $m \geq 1$. We say $\abs{P} = m$ is the \emph{path length}.
+
+\begin{lemma}[Poincare inequality for general graphs.]
+	Define a mapping $\gamma: E_G \to \mathcal{P}$ where for each $e = (\ell,\ell')$ in $E_G$
+	\begin{equation*}
+	\gamma(e) = ((\ell,u), \ldots, (v,\ell'))
+	\end{equation*}
+	meaning $e$ is mapped to a path which begins at $\ell$ and ends at $\ell'$. Then
+	\begin{equation*}
+	G \preceq \widetilde{G} \cdot \max_{e \in E_G} \abs{\gamma(e)}  \cdot b_{\gamma}
+	\end{equation*}
+	where $b_{\gamma}$ is a bottleneck parameter given by
+	\begin{equation*}
+	b_{\gamma} = \max_{\widetilde{e} \in E_{\widetilde{G}}} \abs{\set{e \in E: \widetilde{e} \in \widetilde{P}_e}}
+	\end{equation*}
 \end{lemma}
+\begin{proof}
+	Let $G_e = (V, \set{e})$ and $P_e = (V, \set{\widetilde{e}: \widetilde{e} \in \gamma(e)})$ be the graphs associated with $e$ and $\gamma(e)$, respectively. By Lemma \ref{lem: path_poincare}, we have
+	\begin{equation*}
+	G_{e} \preceq \abs{P_e} P_e
+	\end{equation*}
+	Summing over all $e \in E_G$, we obtain
+	\begin{align*}
+	G & \preceq \sum_{e \in E_G} \abs{P_e} P_e \\
+	& \preceq \max_{e \in E_G} \abs{\gamma(e)} \sum_{e \in E_G} P_e \\
+	& \preceq \max_{e \in E_G} \abs{\gamma(e)} b_{\gamma}\cdot \widetilde{G}
+	\end{align*}
 \end{proof}
+
+\begin{lemma}[Poincare inequality for path graphs.]
+	\label{lem: path_poincare}
+	Fix $m \geq 0$. For vertices $V = \set{1, \ldots,m}$ define the path $P(1 \to m) = ((1,2),(2,3),\ldots, (m-1,m))$ and $G_{(1,m)}$ to be the graph consisting only of an edge between $1$ and $m$. Then,
+	\begin{equation*}
+	(m - 1) \cdot P(1 \to m) \succeq G_{(1,m)}
+	\end{equation*}
+\end{lemma}
 
@@
 
-\clearpage
-
-\bibliography{../../graph_testing_bibliography}
-\bibliographystyle{plain}
-
 \end{document}

### 2019-06-07 notes/weekly_notes/2.29.19.NotesForWeek.tex
weekly notes updated

@@
 \usepackage[shortlabels]{enumitem}
+\usepackage{fullpage}
 
@@
 %%% Matrices
-\newcommand{\Xbf}{\mathbf{X}}
-\newcommand{\Ybf}{\mathbf{Y}}
-\newcommand{\Zbf}{\mathbf{Z}}
+\newcommand{\X}{X} % no bold
+\newcommand{\Y}{Y} % no bold
+\newcommand{\Z}{Z} % no bold
 
@@
 \newcommand{\dive}{\mathrm{div}}
+
+%%% Misc
+\newcommand{\grid}{\mathrm{grid}}
 
@@
 Consider distributions $\Pbb$ and $\Qbb$ supported on $\D \subset \Reals^d$ which are absolutely continuous with density functions $f$ and $g$, respectively.
-For fixed $n \geq 0$, Let $\Zbf = (z_1, \ldots,z_n)$, where for $i = 1,\ldots,t$, $z_i \sim \frac{\Pbb + \Qbb}{2}$ are independent.
-Given $\Zbf$, for $i = 1,...,t$ let
+For fixed $n \geq 0$, let $\Z = (z_1, \ldots,z_n)$, where for $i = 1,\ldots,n$, $z_i \sim \frac{\Pbb + \Qbb}{2}$ are independent.
+Given $\Z$, for $i = 1,...,n$ let
 \begin{equation*}
@@
 \end{equation*} 
-be conditional independent labels, and write
+be conditionally independent labels, and write
 \begin{equation*}
@@
 \end{equation*}
-We will write $\Xbf = \set{x_1, \ldots,x_{N_X}} := \set{z_i: \ell_i = 1}$ and similarly $\Ybf = \set{y_1, \ldots,y_{N_Y}} := \set{y_i: \ell_i = -1}$, where $N_X$ and $N_Y$ are of course random but $N_X + N_Y = n$.
+We will write $\X = \set{x_1, \ldots,x_{N_X}} := \set{z_i: \ell_i = 1}$ and similarly $\Y = \set{y_1, \ldots,y_{N_Y}} := \set{z_i: \ell_i = -1}$, where $N_X$ and $N_Y$ are of course random but $N_X + N_Y = n$.
 
@@
 For $r \geq 0$, define the \emph{$r$-graph} $G_r = (V,E_r)$ to have vertex set $V = \set{1,\ldots,t}$ and edge set $E_r$ which contains the pair $(i,j)$ if and only if $\norm{z_i - z_j}_2 \leq r$.
-Let $D_{G_r}$ denote the incidence matrix of $G_r$.
+Let $D_{r}$ denote the incidence matrix of $G_r$.
 
-Define the \emph{$r$-Laplacian Smooth} test statistic to be
+For a critical radius $C_{n,r}$ to be determined later, define the \emph{$r$-Laplacian Smooth} test statistic to be
 \begin{equation*}
-T_{LS} = \sup_{\theta: \norm{D_{G_r}\theta}_2 \leq C_{n,r}} \dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
+T_{LS} = \sup_{\theta: \norm{D_{r}\theta}_2 \leq C_{n,r}} \dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
 \end{equation*}
@@
 	\label{lem: spectral_similarity_rgraph_grid}
-	Fix $r \geq 2 \left(\frac{\log t}{n}\right)^{1/d} + (\frac{1}{n})^{1/d}$, and let $\ell(r,n) = \biggl(\sqrt{d} r n^{1/d} + 2\sqrt{d}(\log n)^{1/d}\biggr)^{1/2}$.
-For any $\theta \in \Reals^n$, the following relations hold:
+	Fix $r \geq 2 C\left(\frac{\log n}{n}\right)^{1/d} + \left(\frac{1}{n}\right)^{1/d}$, where $C > 0$ is a universal constant, and let 
+	\begin{equation*}
+	\sigma_{r,n} =  d^{d + 1/2} n^{2 + 1/d} \left(2C\left(\frac{\log n}{n}\right)^{1/d} + r \right)^{2d + 1}
+	\end{equation*}
+
+	For any $\theta \in \Reals^n$, there exists a permutation $\Pi: \Reals^d \to \Reals^d$ such that the following relations hold:
 	\begin{equation}
 	\label{eqn: spectral_similarity_rgraph_grid}
-	\frac{\norm{D_{G_r}\theta}_2}{\ell(t)} \leq \norm{D_{grid}\theta}_2 \leq \norm{D_{G_r}\theta}_2
+	\frac{\norm{D_{G_r}\theta}_2}{\sigma_{r,n}} \leq \norm{D_{grid} (\Pi \theta)}_2 \leq \norm{D_{G_r}\theta}_2
 	\end{equation}
@@
 \end{lemma}
-\begin{proof}
-	We begin by mapping the data $\Zbf$ to the grid points $[\kappa]^d$ in such a way that as little mass as possible is disturbed:
+
+Lemma \ref{eqn: spectral_similarity_rgraph_grid} relies heavily on theory regarding optimal transportation matchings between two sets of discrete points, this case $\Z$ and $V_{grid}$.
+
 \begin{lemma}
 	\label{lem: mass_transport_mapping}
-		There exists a bijective mapping $T: \Zbf \to [\kappa]^d$ for $\kappa = t^{1/d}$ such that
+	There exists a bijection $T: \Z \to V_{\grid}$ such that
 	\begin{equation*}
-		\max_{i} \norm{T(z_i) - z_i}_2 \leq C\left(\frac{\log t}{t}\right)^{1/d}
+	\max_{i \in [n]} \norm{T(z_i) - z_i}_2 \leq C\left(\frac{\log n}{n}\right)^{1/d}
 	\end{equation*}
-		with probability at least $1 - n^{-\alpha}$ where $\alpha = c_1 (\log n)^{1/2}$ for some constant $c_1 > 0$.
+	with probability at least $1 - n^{-\alpha}$, where $\alpha = c_1 (\log n)^{1/2}$ and $c_1, C > 0$ are universal constants.
 \end{lemma}
-	Hereafter, we assume there exists $T$ such that Lemma \ref{lem: mass_transport_mapping} holds.
 
-	We first prove the second bound in \eqref{eqn: spectral_similarity_rgraph_grid}.
-Consider grid points $k ~ k'$ connected in the grid graph.
-Then, there exist $z_i$ and $z_j$ such that $T(z_i) = k$ and $T(z_j) = k'$.
+The upper bound of \eqref{eqn: spectral_similarity_rgraph_grid} follows easily.
+\begin{proof}[Upper bound of \eqref{eqn: spectral_similarity_rgraph_grid}]
+	Assume there exists $T$ such that Lemma \ref{lem: mass_transport_mapping} holds.
+
+	Let $k, k' \in [\kappa]^d$ satisfy $\frac{k}{\kappa} ~ \frac{k'}{\kappa}$  in the grid graph.
+There exist $z_i$ and $z_j$ such that $T(z_i) = \frac{k}{\kappa}$ and $T(z_j) = \frac{k'}{\kappa}$.
 By the triangle inequality,
@@
 	\norm{z_i - z_j}_2 & \leq \norm{T(z_i) - z_i}_2 + \norm{T(z_i) - T(z_j)}_2 + \norm{T(z_j) - z_j}_2 \\
-	& \leq 2C\left(\frac{\log t}{t}\right)^{1/d} + \frac{1}{t^{1/d}}
+	& \leq 2C\left(\frac{\log n}{n}\right)^{1/d} + \frac{1}{n^{1/d}}
 	\end{align*}
 	and so by our choice of $r$, $i \sim j$ in $G_r$.
+\end{proof}
 
-	Now, we turn to the first bound. Assume $i \sim j$ in the graph $G_r$. By a similar set of steps to the above, we have
+To show the lower bound of \eqref{eqn: spectral_similarity_rgraph_grid}, we will make use of a technique from spectral graph theory known as Poincare's inequality.
+
+\paragraph{Poincare inequality}
+
+Let $G$ and $\widetilde{G}$ be undirected, unweighted graphs over vertex set $V$, with edge sets $E_G$ and $E_{\widetilde{G}}$, respectively.
+Let $\widetilde{\mathcal{P}}$ be the space of all paths over $E_{\widetilde{G}}$; that is, $\mathcal{P}$ consists of $\widetilde{P} \in \widetilde{\mathcal{P}}$ with
+\begin{equation*}
+\widetilde{P} = (\widetilde{e}_1, \ldots, \widetilde{e}_m) \tag{$\widetilde{e_i} \in E_{\widetilde{G}}$}
+\end{equation*}
+for some integer $m \geq 1$.
+
+\begin{lemma}[Poincare inequality]
+	\label{lem: poincare}
+	Define a mapping $\gamma: E_G \to \mathcal{P}$ where for each $e = (\ell,\ell')$ in $E_G$
+	\begin{equation*}
+	\gamma(e) = ((\ell,u), \ldots, (v,\ell'))
+	\end{equation*}
+	meaning $e$ is mapped to a path which begins at $\ell$ and ends at $\ell'$. Then
+	\begin{equation*}
+	G \preceq \widetilde{G} \cdot \max_{e \in E_G} \abs{\gamma(e)}  \cdot b_{\gamma}
+	\end{equation*}
+	where $b_{\gamma}$ is a bottleneck parameter given by
+	\begin{equation*}
+	b_{\gamma} = \max_{\widetilde{e} \in E_{\widetilde{G}}} \abs{\set{e \in E: \widetilde{e} \in \gamma(e)}}
+	\end{equation*}
+\end{lemma}
+
+Lemma \ref{lem: mass_transport_mapping} will allow us to construct such a mapping $\gamma$ from $E_r$ to $E_{\grid}$ and appropriately control parameters $\max_{e \in E_G} \abs{\gamma(e)}$ and $b_{\gamma}$.
+
+\begin{lemma}
+	There exists a mapping $\gamma: E_r \to \mathcal{P}_{\grid}$, the set of paths over $G_{\grid}$, such that the following quantities are bounded:
+	\begin{enumerate}[(i)]
+		\item \textit{Maximum path length.} 
+		\begin{equation*}
+		\max_{e \in E_G} \abs{\gamma(e)} \leq n^{1/d} \sqrt{d} \left(2C\left(\frac{\log n}{n}\right)^{1/d} + r \right)
+		\end{equation*}
+		\item \textit{Bottleneck.}
+		\begin{equation*}
+		b_{\gamma} \leq \left(n^{1/d} \sqrt{d} \left(2C\left(\frac{\log n}{n}\right)^{1/d} + r \right) \right)^{2d}
+		\end{equation*}
+	\end{enumerate}
+	with probability at least $1 - n^{-\alpha}$ where $\alpha = c_1 (\log n)^{1/2}$ and $C, c_1 > 0$ are universal constants.
+\end{lemma}
+
+\begin{proof}	
+	Assume $i \sim j$ in the graph $G_r$. By a similar set of steps to the above, we have
 	\begin{equation*}
@@
 
-\section{Poincare inequalities}
+\section{Additional Theory and Proofs}
 
-Let $G$ and $\widetilde{G}$ be undirected, unweighted graphs over vertex set $V$, with edge sets $E_G$ and $E_{\widetilde{G}}$, respectively.
-Let $\widetilde{\mathcal{P}}$ be the space of all paths over $E_{\widetilde{G}}$; that is, $\mathcal{P}$ consists of $\widetilde{P} \in \widetilde{\mathcal{P}}$
+\subsection{Proof of Lemma \ref{lem: poincare}}
+
+\begin{lemma}[Poincare inequality for path graphs.]
+	\label{lem: path_poincare}
+	Fix $m \geq 0$. For vertices $V = \set{1, \ldots,m}$ define the path $P(1 \to m) = ((1,2),(2,3),\ldots, (m-1,m))$ and $G_{(1,m)}$ to be the graph consisting only of an edge between $1$ and $m$. Then,
 	\begin{equation*}
-P = (\widetilde{e}_1, \ldots, \widetilde{e}_m) \tag{$\widetilde{e_i} \in E_{\widetilde{G}}$}
-\end{equation*}
-for some integer $m \geq 1$. We say $\abs{P} = m$ is the \emph{path length}.
-
-\begin{lemma}[Poincare inequality for general graphs.]
-	Define a mapping $\gamma: E_G \to \mathcal{P}$ where for each $e = (\ell,\ell')$ in $E_G$
-	\begin{equation*}
-	\gamma(e) = ((\ell,u), \ldots, (v,\ell'))
-	\end{equation*}
-	meaning $e$ is mapped to a path which begins at $\ell$ and ends at $\ell'$. Then
-	\begin{equation*}
-	G \preceq \widetilde{G} \cdot \max_{e \in E_G} \abs{\gamma(e)}  \cdot b_{\gamma}
-	\end{equation*}
-	where $b_{\gamma}$ is a bottleneck parameter given by
-	\begin{equation*}
-	b_{\gamma} = \max_{\widetilde{e} \in E_{\widetilde{G}}} \abs{\set{e \in E: \widetilde{e} \in \widetilde{P}_e}}
+	(m - 1) \cdot P(1 \to m) \succeq G_{(1,m)}
 	\end{equation*}
 \end{lemma}
-\begin{proof}
+
+\paragraph{Proof of Lemma \ref{lem: poincare}}
 Let $G_e = (V, \set{e})$ and $P_e = (V, \set{\widetilde{e}: \widetilde{e} \in \gamma(e)})$ be the graphs associated with $e$ and $\gamma(e)$, respectively. By Lemma \ref{lem: path_poincare}, we have
@@
 \end{align*}
-\end{proof}
-
-\begin{lemma}[Poincare inequality for path graphs.]
-	\label{lem: path_poincare}
-	Fix $m \geq 0$. For vertices $V = \set{1, \ldots,m}$ define the path $P(1 \to m) = ((1,2),(2,3),\ldots, (m-1,m))$ and $G_{(1,m)}$ to be the graph consisting only of an edge between $1$ and $m$. Then,
-	\begin{equation*}
-	(m - 1) \cdot P(1 \to m) \succeq G_{(1,m)}
-	\end{equation*}
-\end{lemma}
 

### 2019-06-07 notes/weekly_notes/5.15.19.NotesForWeek.tex
weekly notes updated

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbb{I}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 5/15/19 - 5/17/19}
\author{Alden Green}
\date{\today}
\maketitle

Consider absolutely continuous distributions $\Pbb$ and $\Qbb$ with density functions $f$ and $g$, respectively. For fixed $n \geq 0$, let $\Z = (z_1, \ldots,z_n)$, where for $i = 1,\ldots,n$, $z_i \sim \frac{\Pbb + \Qbb}{2}$ are independent. Given $\Z$, for $i = 1,...,n$ let
\begin{equation*}
\ell_i = 
\begin{cases}
1~ \text{with probability $\frac{f(z_i)}{f(z_i) + g(z_i)}$} \\
-1~ \text{with probability $\frac{g(z_i)}{f(z_i) + g(z_i)}$}
\end{cases}
\end{equation*} 
be conditionally independent labels, and write
\begin{equation*}
1_X = 
\begin{cases}
1,~ l_i = 1\\
0,~ \text{otherwise}
\end{cases}
1_Y = 
\begin{cases}
1,~ l_i = -1 \\
0,~ \text{otherwise.}
\end{cases}
\end{equation*}
We will write $\X = \set{x_1, \ldots,x_{N_X}} := \set{z_i: \ell_i = 1}$ and similarly $\Y = \set{y_1, \ldots,y_{N_Y}} := \set{z_i: \ell_i = -1}$, where $N_X$ and $N_Y$ are of course random but $N_X + N_Y = n$.

Our statistical goal is hypothesis testing: that is, we wish to construct a test function $\phi$ which differentiates between
\begin{equation*}
\mathbb{H}_0: f = g \text{ and } \mathbb{H}_1: f \neq g.
\end{equation*}

For a given function class $\Hclass$, some $\epsilon > 0$, and test function $\phi$ a Borel measurable function of the data with range $\set{0,1}$, we evaluate the quality of the test using \emph{worst-case risk}
\begin{equation*}
R_{\epsilon}^{(t)}(\phi; \Hclass) = \sup_{f \in \Hclass} \Ebb_{f,f}^{(t)}(\phi) + \sup_{ \substack{f,g \in \Hclass \\ \delta(f,g) \geq \epsilon } } \Ebb_{f,g}^{(t)}(1 - \phi)
\end{equation*} 
where 
\begin{equation*}
\delta^2(f,g) = \int_{\D} (f - g)^2 dx.
\end{equation*}

\paragraph{Test statistic.}

For $r \geq 0$, define the \emph{$r$-graph} $G_r = (V,E_r)$ to have vertex set $V = \set{1,\ldots,t}$ and edge set $E_r$ which contains the pair $(i,j)$ if and only if $\norm{z_i - z_j}_2 \leq r$. Let $D_{r}$ denote the incidence matrix of $G_r$.

Define the \emph{Laplacian Smooth} test statistic over the neighborhood graph to be
\begin{equation*}
T_{LS} = \sup_{\theta: \norm{D_{r}\theta}_2 \leq C(n,r)} \dotp{\theta }{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
\end{equation*}
where we note that the test statistic is implicitly a function of $r$ and $C(n,r)$.

\paragraph{General assumptions.}
We will need to make the following assumptions in order to obtain meaningful theoretical results.

\begin{assumption}[General assumptions.]
	\label{asmp: general}
	For each $f \in \mathcal{H}$, assume $\mathrm{supp}(f)$ is an open connected, bounded subset of $\Rd$ with $l$-Lipschitz boundary. Further assume $\mathcal{H}$ is uniformly bounded above and below: that is, there exists $f_{\min} < f_{\max}$ such that for each $f \in \mathcal{H}$, 
	\begin{equation*}
	0 < f_{\min} < f(x) < f_{\max} < \infty, \quad \textrm{for all $x \in \mathrm{supp}(f)$}
	\end{equation*}
\end{assumption}

We motivate these assumptions through the following theorem relating the $r$-neighborhood graph to the $d$-dimensional grid, a useful result as the spectrum of grid graphs is well known.

Formally, for $\kappa = n^{1/d}$, consider the \emph{grid graph}
\begin{equation*}
G_{grid} = (V_{grid},E_{grid}),~~ V_{grid} = \set{k: k \in [\kappa]^d},~~ E_{grid} = \set{(k,k'): k, k' \in V_{grid}, \norm{k - k'}_1 = 1}
\end{equation*}
with associated incidence matrix $D_{grid}$.

\begin{lemma}
	Let $z_1, \ldots,z_n \sim p$ for some density function $p \in \mathcal{H}$ which satisfies Assumption \ref{asmp: general}.

	Fix $a > 2$. There exists $c_1 > 0$ (potentially depending on $d, l, f_{\min}, f_{\max}$ and $a$) such that for any $r \geq c_1 \left(\frac{\log n}{n}\right)^{1/d}$, the following statement holds with probability at least $1 - n^{-a}$: there exists a bijection $T: [n] \to [\kappa]^d$ such that for all $\theta = (\theta_1, \ldots, \theta_n) \in \Reals^n$, letting
	\begin{equation*}
	(\theta_{T})_{k} = \theta_{T^{-1}(k)}, \textrm{for all $k \in [\kappa]^d$}
	\end{equation*}
	we have
	\begin{equation*}
	\norm{D_{\grid}\theta_{T}}_2 \leq \norm{D_r \theta}_2.
	\end{equation*}
\end{lemma}
\textcolor{red}{Prove this.}

We use this to control the empirical process $T_{LS}$ under the null hypothesis.
\begin{lemma}
	Assume $\mathcal{H}$ satisfies Assumption \ref{asmp: general}, and fix $a \geq 2$. Then, there exists $c_1 > 0$ (potentially depending on $d, l, f_{\min}$, $f_{\max}$, and $a$) and $c_2$ (potentially depending on $d$) such that for any $r \geq c_1 \left(\frac{\log n}{n}\right)^{1/d}$, and any $C_{n,r} > 0$,
	\begin{equation*}
	\Pbb_{\mathbb{H}_0}\left(T_{LS} \leq \frac{c_2 C(n,r)\sqrt{\log n} }{a \sqrt{n}} \right) \leq 1 - 2\exp(2a) - a
	\end{equation*}
\end{lemma}
\begin{proof}
	We write
	\begin{align*}
	T_{LS} & \leq \abs{T_{LS} - \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y}} + \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y} \\
	& \leq \abs{T_{LS} - \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y}} + \sup_{\substack{\theta: \norm{D_{\grid} \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y} \tag{Lemma \ref{lem: closed_form_ls_statistic} } \\
	& \leq \abs{T_{LS} - \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y}} + C(n,r) \frac{1}{n/2} \sqrt{\sigma^T L_{\grid}^{\dagger} \sigma}
	\end{align*}
	whence the statement follows by Lemmas \ref{lem: normalization_error} and \ref{lem: noise_quadratic_form_on_grid}.
\end{proof}

\section{Additional Theory}

The Laplacian smooth test statistic has a closed form solution. Let $L = D^T D$ be the Laplacian of $G$, and write $\Linv$ for the pseudoinverse of $L$.

\begin{lemma}
	\label{lem: closed_form_ls_statistic}
	For any unweighted, undirected, connected graph $G = (V,E)$ with incidence matrix $D$, number $C > 0$, and any vector $v \in \Reals^n$ with $\sum_{i = 1}^{n} v_i = 0$,
	\begin{equation*}
	\sup_{\theta: \norm{D\theta}_2 \leq C} \dotp{\theta}{v} = C \sqrt{v^T \Linv v}
	\end{equation*}
	Additionally, for any vector $v \in \Reals^n$ (not necessarily $\sum_{i = 1}^{n} v_i = 0$), under the additional constraint $\theta^T \mathbf{1} = 0$, the same statement holds. That is,
	\begin{equation*}
	\sup_{ \substack{\theta: \norm{D\theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \dotp{\theta}{v} = C \sqrt{v^T \Linv v}
	\end{equation*}
\end{lemma}
\begin{proof}
	Note that the condition $\norm{D \theta}_2 \leq C$ is equivalent to $\theta^T L \theta \leq C$. The solution then follows from the KKT conditions.
\end{proof}

\begin{lemma}
	\label{lem: normalization_error}
	Under the general assumptions,
	\begin{equation*}
	\abs{T_{LS} - \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y}} \leq \frac{8C(n,r)}{a n^{1 - 1/d}\pi}
	\end{equation*}
	with probability at least $1 - 2 \exp(2a)$. 
\end{lemma}
\begin{proof}
	Note that as $(1_X/N_X - 1_Y/X_Y)^T \mathbf{1} = 0$, we may write
	\begin{align*}
	T_{LS} & = \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq 1, \\ \theta^T \mathbf{1} = 0} } \dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
	\end{align*}
	and therefore by Cauchy-Schwarz,
	\begin{align*}
	\abs{T_{LS} - \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y}}  & \leq \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \cdot \norm{1_X(2/n - 1/N_X) - 1_Y(1/N_Y - 2/n)} \\
	& \leq \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \cdot \norm{1_X(2/n - 1/N_X) - 1_Y(1/N_Y - 2/n)} \\
	& \leq \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \cdot \left(\norm{1_X} \frac{\abs{n/2 - N_X}}{n/2 N_X} + \norm{1_Y} \frac{\abs{n/2 - N_Y}}{n/2 N_Y}\right) \\
	& \leq \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \cdot \left(\frac{2\abs{n/2 - N_X}}{\sqrt{N_X} n} + \frac{2\abs{n/2 - N_Y}}{\sqrt{N_Y} n} \right).
	\end{align*}
	The reasoning in the proof of Lemma \ref{lem: noise_quadratic_form_on_grid} leads to the bound
	\begin{equation*}
	\sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \leq \frac{C n^{1/d}}{\pi},
	\end{equation*}
	and a standard application of Hoeffding's inequality to the quantities $N_X$ and $N_Y$ yields, for any $a > 0$
	\begin{equation*}
	\abs{N_X - n/2}, \abs{N_Y - n/2} \leq  a \sqrt{n}
	\end{equation*}
	with probability at least $1 - 2 \exp(-2a)$. As a result,
	\begin{equation*}
	\sup_{\substack{\theta: \norm{D_r \theta}_2 \leq 1, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \cdot \left(\frac{2\abs{n/2 - N_X}}{\sqrt{N_X} n} + \frac{2\abs{n/2 - N_Y}}{\sqrt{N_Y} n} \right) \leq \frac{8a C n^{1/d}}{n\pi} 
	\end{equation*}
	with probability at least $1 - 2 \exp(-2a)$, which proves the claim.
\end{proof}

\subsection{Type I error.}

Denote the eigenvalues and eigenvectors of $\Lgrid = \Dgrid^T \Dgrid$ as $\set{\lambda_{k}: k \in [\kappa]^d}$ and $\set{u_{k}: k \in [\kappa]^d}$, respectively.

\begin{lemma}
	\label{lem: noise_quadratic_form_on_grid}
	Let $\sigma_1, \ldots, \sigma_n \sim \mathrm{Rademacher}(1/2)$ be independent random variables, and write $\sigma = (\sigma_1, \ldots, \sigma_n)$. Then, 
	\begin{equation*}
	\sigma^T \Linvgrid \sigma \leq \frac{(2d)^d\kappa^2}{a\pi^2} \left(8 + 2\pi \log(\sqrt{2} \kappa) \left(\sum_{p = 2}^{d}\kappa^{p}\right) \right)
	\end{equation*}
	with probability at least $1  - a$.
\end{lemma}
\begin{proof}
	Taking the eigendecomposition $\Lgrid = U \Lambda U^T$, we can then write $\Linvgrid = U \Lambda^{\dagger} U^T$. Therefore
	\begin{equation*}
	\Ebb \bigl(\sigma^T \Linvgrid \sigma\bigr) = \sum_{k \in [\kappa]^d} \frac{\Ebb \bigl(u_{k}^T \sigma\bigr)^2}{\lambda_k}
	\end{equation*}
	and as 
	\begin{equation*}
	\Ebb \bigl(u_{k}^T \sigma\bigr)^2 = \sum_{i,j = 1}^{n} \Ebb(u_{k_i} u_{k_j} \sigma_i \sigma_j) = \sum_{i = 1}^{n} u_{k_i}^2 = 1
	\end{equation*}
	we are left with
	\begin{equation}
	\label{eqn: noise_quadratic_form_on_grid_1}
	\Ebb \bigl(\sigma^T \Linvgrid \sigma\bigr) = \sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \frac{1}{\lambda_k}.
	\end{equation}

	It is well known \textcolor{red}{HR, CB}, that the $d$-dimensional grid can be written as the Kronecker product of $d$ path graphs, and exploiting this fact as in \textcolor{red}{SWT} we obtain
	\begin{align*}
	\lambda_{k} = 4 \sin^2\left(\frac{\pi(k_1 - 1)}{2 \kappa}\right) + \cdots + 4 \sin^2\left(\frac{\pi(k_d - 1)}{2 \kappa}\right).
	\end{align*}
	The inequality $\sin(x) \geq x/2$ holds for $x \in [0,\pi/2]$, and so we have
	\begin{align*}
	\sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \frac{1}{\lambda_k} & \leq \frac{\kappa^2}{\pi^2} \sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d} (k_j - 1)^2\right)^{-1} \\
	& \leq \frac{(2d)^d\kappa^2}{\pi^2} \left(8 + 2\pi \log(\sqrt{2} \kappa) \left(\sum_{p = 2}^{d}\kappa^{p - 2}\right) \right) \\
	& = \frac{(2d)^d\kappa^2}{\pi^2} \left(8 + 2\pi \log(\sqrt{2} \kappa) \left(\sum_{p = 2}^{d}\kappa^{p}\right) \right) \tag{Lemma \ref{lem: grid_eigenvalues}}
	\end{align*}
	whence the claim follows by Markov's inequality.
\end{proof}

\begin{lemma}
	\label{lem: grid_eigenvalues}
	There exists a universal constant $c_1 > 0$ such that for any $\kappa \geq 1$ and for all integers $d \geq 3$,
	\begin{equation*}
	\sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d} (k_j - 1)^2\right)^{-1} \leq (2d)^d \Biggl(8 + 2 \pi \log (\sqrt{2} \kappa) \biggl(\sum_{p = 2}^{d} \kappa^{p - 2}\biggr) \Biggr)
	\end{equation*}
\end{lemma}
\begin{proof}
	We will prove by induction on $d$. As a base case, consider $d  = 2$. Rewrite
	\begin{equation*}
	\sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d} (k_j - 1)^2\right)^{-1} \leq 4 \sum_{k_1 = 1}^{\kappa - 1} \frac{1}{k_1^2}  + \sum_{k_1 =2}^{\kappa - 1} \sum_{k_2 = 2}^{\kappa - 1} (k_1^2 + k_2^2)^{-1}.
	\end{equation*}
	Then, upper bounding sums by integrals, we obtain
	\begin{equation*}
	\sum_{k_1 = 1}^{\kappa - 1} \frac{1}{k_1^2} \leq 1 +  \int_{x = 1}^{\kappa} \frac{1}{x^2} \dif x \leq 2,
	\end{equation*}
	and 
	\begin{align*}
	\sum_{k_1 = 2}^{\kappa - 1} \sum_{k_2 = 2}^{\kappa - 1} (k_1^2 + k_2^2)^{-1} & \leq \int_{x = 1}^{\kappa} \int_{y = 1}^{\kappa} \frac{1}{x^2 + y^2} \dif x \dif y \\
	& \leq 2 \pi \log(\sqrt{2} \kappa),
	\end{align*}
	so that
	\begin{equation*}
	\sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d} (k_j - 1)^2\right)^{-1} \leq 8 + 2 \pi \log(\sqrt{2} \kappa).
	\end{equation*}
	and the base case is shown.

	We proceed by induction. For $d \geq 3$ we write
	\begin{align*}
	\sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d} (k_j - 1)^2\right)^{-1} & \leq 2d \sum_{\substack{k \in [\kappa]^{d - 1}, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d - 1} (k_j - 1)^2\right)^{-1} + \sum_{k_1 = 2}^{\kappa - 1} \cdots \sum_{k_d = 2}^{\kappa - 1} \left(\sum_{j = 1}^{d} k_j^2\right)^{-1} \\
	& \leq 2d \left( (2d)^{d - 1}\Biggl(8 + 2 \pi \log \sqrt{2} \kappa \biggl(\sum_{p = 2}^{d - 1} \kappa^{p - 2}\biggr) \Biggr)\right) + \sum_{k_1 = 2}^{\kappa - 1} \cdots \sum_{k_d = 2}^{\kappa - 1} \left(\sum_{j = 1}^{d} k_j^2\right)^{-1} \\
	& = \left( (2d)^{d}\Biggl(8 + 2 \pi \log \sqrt{2} \kappa \biggl(\sum_{p = 2}^{d - 1} \kappa^{p - 2}\biggr) \Biggr)\right) + \sum_{k_1 = 2}^{\kappa - 1} \cdots \sum_{k_d = 2}^{\kappa - 1} \left(\sum_{j = 1}^{d} k_j^2\right)^{-1} 
	\end{align*}
	and again upper bounding sums by integrals, we obtain
	\begin{align*}
	\sum_{k_1 = 2}^{\kappa - 1} \cdots \sum_{k_d = 2}^{\kappa - 1} \left(\sum_{j = 1}^{d} (k_j)^2\right)^{-1} & \leq \int_{1}^{\kappa} \cdots \int_{1}^{\kappa} \frac{1}{x_1^2 + \cdots + x_d^2} \dif x_d \ldots \dif x_1 \\
	& \leq \kappa^{d-2} \int_{1}^{\kappa} \int_{1}^{\kappa} \frac{1}{x^2 + y^2} \dif x \dif y \\
	& \leq \kappa^{d-2} 2 \pi \log(\sqrt{2} \kappa)
	\end{align*}
	and the proof is complete.
\end{proof}

\paragraph{Type II error.}

Lemma \ref{lem: approx_error_lb_1} will be useful in lower bounding the \textcolor{red}{approximation error} of the $T_{LS}$ test statistic. Write $P_1^{\perp}$ for the projection operator onto the subspace of $\Reals^n$ orthogonal to the constant vector $\mathbf{1}$.

\begin{lemma}
	\label{lem: approx_error_lb_1}
	For any vector $v \in \Reals^n$ and Laplacian matrix $L$ of a connected graph $G$,
	\begin{equation*}
	v^T \Linv v \geq \frac{(v^T P_{1}^{\perp}v)^2}{v^T L v}
	\end{equation*}
\end{lemma}
\begin{proof}
	We can expand $v^T P_1^{\perp} v  = v^T (L)^{1/2}(\Linv)^{1/2} v$, for any matrix square root of $L$ and $\Linv$. The statement follows by Cauchy-Schwarz.
\end{proof}

\end{document}

### 2019-06-07 notes/weekly_notes/5.29.19.NotesForWeek.tex
weekly notes updated

--- 2019-06-07 notes/weekly_notes/5.15.19.NotesForWeek.tex
@@
 %%% Vectors
+\newcommand{\thetast}{\theta^{\star}}
 
@@
 \newcommand{\dif}{\mathop{}\!\mathrm{d}}
+\newcommand{\gradient}{\mathcal{D}}
+\newcommand{\Hessian}{\mathcal{D}^2}
 
@@
 \newcommand{\critr}{R_n}
+\newcommand{\dx}{\,dx}
+\newcommand{\dy}{\,dy}
 
@@
 \begin{document}
-\title{Notes for Week 5/15/19 - 5/17/19}
+\title{Notes for Week 5/29/19 - 5/31/19}
 \author{Alden Green}
@@
 \begin{equation*}
-T_{LS} = \sup_{\theta: \norm{D_{r}\theta}_2 \leq C(n,r)} \dotp{\theta }{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
+T_{LS} = \sup_{\theta: \norm{D\theta}_2 \leq C(n,r)} \dotp{\theta }{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
 \end{equation*}
@@
 
-\paragraph{General assumptions.}
-We will need to make the following assumptions in order to obtain meaningful theoretical results.
-
-\begin{assumption}[General assumptions.]
-	\label{asmp: general}
-	For each $f \in \mathcal{H}$, assume $\mathrm{supp}(f)$ is an open connected, bounded subset of $\Rd$ with $l$-Lipschitz boundary.
-Further assume $\mathcal{H}$ is uniformly bounded above and below: that is, there exists $f_{\min} < f_{\max}$ such that for each $f \in \mathcal{H}$,
+\section{Additional Theory}
+Let $\theta^{\star} = (\thetast_i)_{i = 1}^n$, with $\thetast_i := f(z_i) - g(z_i)$.
+Write $L = D^T D$ for the Laplacian matrix of the $r$-neighborhood graph, $\gradient f$ for the gradient of a function $f$, and $\Hessian f$ for the Hessian of a function $f$.
+We write $C^{2}(L)$ for the set of functions $f$ twice continuously differentiable over $\Rd$, with bounded Hessian $\norm{\Hessian f}_{\infty} \leq L$.
+\begin{lemma}
+	\label{lem: approx_error}
+	For all density functions $f,g \in C^2(L)$ with $\int (f - g)^2 \geq \epsilon^2$, if $r \geq c_1 (\log n / n)^{1/d}$ and $\epsilon \geq c_2 n r^{d + 1} $ then
 	\begin{equation*}
-	0 < f_{\min} < f(x) < f_{\max} < \infty, \quad \textrm{for all $x \in \mathrm{supp}(f)$}
+	\sup_{\norm{D\theta}_2 \leq 1} \dotp{\theta}{\thetast} \geq \frac{c_2}{2} (\log n)^{1 + 2/d} n^{-2/d}
 	\end{equation*}
-\end{assumption}
-
-We motivate these assumptions through the following theorem relating the $r$-neighborhood graph to the $d$-dimensional grid, a useful result as the spectrum of grid graphs is well known.
-
-Formally, for $\kappa = n^{1/d}$, consider the \emph{grid graph}
+	with probability at least $1 - \delta$.
+\end{lemma}
+\begin{proof}
+	By Lemmas \ref{lem: closed_form_ls_statistic} and \ref{lem: lin_alg_1}, we have
+	\begin{align}
+	\label{eqn: approx_error_1}
+	\sup_{\norm{D\theta}_2 \leq 1} \dotp{\theta}{\thetast} & = \frac{1}{n} \sqrt{(\thetast)^T \Linv \thetast} \nonumber \\
+	& \geq \frac{1}{\lambda_k}\left(\dotp{\thetast}{\thetast} - \dotp{P_k^{\perp} \thetast}{P_k^{\perp}\thetast}\right)^2 
+	\end{align}
+	where the latter inequality holds for any $k = 1, \ldots, n - 1$. We upper bound $\dotp{P_k^{\perp} \thetast}{P_k^{\perp}\thetast}$ using the following relations
 	\begin{equation*}
-G_{grid} = (V_{grid},E_{grid}),~~ V_{grid} = \set{k: k \in [\kappa]^d},~~ E_{grid} = \set{(k,k'): k, k' \in V_{grid}, \norm{k - k'}_1 = 1}
+	(\thetast)^T L \thetast \geq (P_k^{\perp}\thetast)^T \Linv ( P_k^{\perp}\thetast) \geq \lambda_k \dotp{P_k^{\perp} \thetast}{P_k^{\perp} \thetast}
 	\end{equation*}
-with associated incidence matrix $D_{grid}$.
+	and obtain
+	\begin{align*}
+	\frac{1}{\lambda_k}\left(\dotp{\thetast}{\thetast} - \dotp{P_k^{\perp} \thetast}{P_k^{\perp}\thetast}\right)^2  & \geq \frac{1}{\lambda_k}\left(\dotp{\thetast}{\thetast} - \frac{(\thetast)^T L \thetast}{\lambda_k} \right)^2 \\
+	& \geq \frac{1}{\lambda_k}\left(c_1 n \epsilon^2 - \frac{c_2 n^2 r^{d + 2}}{\lambda_k} \right)^2 \tag{Lemmas \ref{lem: witness_inner_product} and \ref{lem: witness_discrete_energy}}
+	\end{align*}
+	where the latter inequality occurs with probability at least $1 - 2\delta$.
+Choose $k = n - 1$.
+As $r \geq c_1 (\log n / n)^{1/d}$, we have that $G \preceq \mathrm{Grid}$, and as a result \textcolor{red}{$4 \geq \lambda_{n - 1} \geq 1$}.
+Therefore,
+	\begin{align*}
+	\frac{1}{\lambda_k}\left(c_1 n \epsilon^2 - \frac{c_2 n^2 r^{d + 2}}{\lambda_k} \right)^2 & \geq \frac{1}{4} \left(c_1 n \epsilon^2 - c_2 n^2 r^{d + 2} \right)^2 \\
+	& \geq \frac{1}{4} \left(c_2 n^2 r^{d + 2}\right)^2 \\
+	& \geq \frac{1}{4} \left(c_2 n^{1 - 2/d} \log(n)^{1 + 2/d}\right)^2 = \frac{c_2^2}{4} n^{2 - 4/d} \log(n)^{2 + 4/d},
+	\end{align*}
+	and the proof is complete.
+\end{proof}
 
 \begin{lemma}
-	Let $z_1, \ldots,z_n \sim p$ for some density function $p \in \mathcal{H}$ which satisfies Assumption \ref{asmp: general}.
+	\label{lem: witness_inner_product}
+	For any $f, g \in L^2(\Rd)$ satisfying the \textcolor{red}{regularity conditions}, such that $\int_{\Rd} (f - g)^2 dx \geq \epsilon^2$ there exists constant $c_1$ such that 
+	\begin{equation*}
+	\dotp{\thetast}{\thetast} \geq c_1 n \epsilon^2
+	\end{equation*}
+	with probability at least $1 - \delta$.
+\end{lemma}
 
-	Fix $a > 2$.
-There exists $c_1 > 0$ (potentially depending on $d, l, f_{\min}, f_{\max}$ and $a$) such that for any $r \geq c_1 \left(\frac{\log n}{n}\right)^{1/d}$, the following statement holds with probability at least $1 - n^{-a}$: there exists a bijection $T: [n] \to [\kappa]^d$ such that for all $\theta = (\theta_1, \ldots, \theta_n) \in \Reals^n$, letting
+\begin{lemma}
+	\label{lem: witness_discrete_energy}
+	For any $f, g \in C^2(L)$ satisfying the \textcolor{red}{regularity conditions}, there exists a constant $c_2$ such that
 	\begin{equation*}
-	(\theta_{T})_{k} = \theta_{T^{-1}(k)}, \textrm{for all $k \in [\kappa]^d$}
+	(\thetast)^T L \thetast \leq c_2 n^2 r^{d + 2}
 	\end{equation*}
-	we have
-	\begin{equation*}
-	\norm{D_{\grid}\theta_{T}}_2 \leq \norm{D_r \theta}_2.
-	\end{equation*}
+	with probability at least $1 - \delta$.
 \end{lemma}
-\textcolor{red}{Prove this.}
 
-We use this to control the empirical process $T_{LS}$ under the null hypothesis.
-\begin{lemma}
-	Assume $\mathcal{H}$ satisfies Assumption \ref{asmp: general}, and fix $a \geq 2$.
-Then, there exists $c_1 > 0$ (potentially depending on $d, l, f_{\min}$, $f_{\max}$, and $a$) and $c_2$ (potentially depending on $d$) such that for any $r \geq c_1 \left(\frac{\log n}{n}\right)^{1/d}$, and any $C_{n,r} > 0$,
-	\begin{equation*}
-	\Pbb_{\mathbb{H}_0}\left(T_{LS} \leq \frac{c_2 C(n,r)\sqrt{\log n} }{a \sqrt{n}} \right) \leq 1 - 2\exp(2a) - a
-	\end{equation*}
-\end{lemma}
-\begin{proof}
-	We write
-	\begin{align*}
-	T_{LS} & \leq \abs{T_{LS} - \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y}} + \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y} \\
-	& \leq \abs{T_{LS} - \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y}} + \sup_{\substack{\theta: \norm{D_{\grid} \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y} \tag{Lemma \ref{lem: closed_form_ls_statistic} } \\
-	& \leq \abs{T_{LS} - \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y}} + C(n,r) \frac{1}{n/2} \sqrt{\sigma^T L_{\grid}^{\dagger} \sigma}
-	\end{align*}
-	whence the statement follows by Lemmas \ref{lem: normalization_error} and \ref{lem: noise_quadratic_form_on_grid}.
-\end{proof}
-
-\section{Additional Theory}
-
-The Laplacian smooth test statistic has a closed form solution. Let $L = D^T D$ be the Laplacian of $G$, and write $\Linv$ for the pseudoinverse of $L$.
+\section{Linear Algebra}
 
@@
 	\label{lem: closed_form_ls_statistic}
-	For any unweighted, undirected, connected graph $G = (V,E)$ with incidence matrix $D$, number $C > 0$, and any vector $v \in \Reals^n$ with $\sum_{i = 1}^{n} v_i = 0$,
+	For any unweighted, undirected, connected graph $G = (V,E)$ with incidence matrix $D$, and any vector $v \in \Reals^n$ with $\sum_{i = 1}^{n} v_i = 0$,
 	\begin{equation*}
@@
 
+Let $L$ be the Laplacian matrix of a connected graph, and $\Linv$ be the pseudo-inverse $L$.
+Write the eigendecomposition of $\Linv = U\Lambda^{\dagger}U^{T}$, where $\Lambda$ is an $n \times n$ diagonal matrix with entries $0 = \lambda_0 < \lambda_1 \leq \lambda_2 \leq \ldots \leq \lambda_{n-1}$ and $U$ is an orthogonal matrix with columns $U =\left(u_0 \ldots u_{n-1}\right)$.
+For each $k = 0, \ldots, n - 1$, write $U_{k} = (u_0 \ldots u_{k-1})$ for the first $k$ columns of $U$, $P_k$ as the projection operator onto the span of $U_k$, and $P_k^{\perp}$ for the projection operator onto the subspace orthogonal to the span of $U_k$.
 \begin{lemma}
-	\label{lem: normalization_error}
-	Under the general assumptions,
+	\label{lem: lin_alg_1}
+	For any $k = 1, \ldots, n$, 
 	\begin{equation*}
-	\abs{T_{LS} - \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y}} \leq \frac{8C(n,r)}{a n^{1 - 1/d}\pi}
-	\end{equation*}
-	with probability at least $1 - 2 \exp(2a)$. 
-\end{lemma}
-\begin{proof}
-	Note that as $(1_X/N_X - 1_Y/X_Y)^T \mathbf{1} = 0$, we may write
-	\begin{align*}
-	T_{LS} & = \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq 1, \\ \theta^T \mathbf{1} = 0} } \dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
-	\end{align*}
-	and therefore by Cauchy-Schwarz,
-	\begin{align*}
-	\abs{T_{LS} - \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y}}  & \leq \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \cdot \norm{1_X(2/n - 1/N_X) - 1_Y(1/N_Y - 2/n)} \\
-	& \leq \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \cdot \norm{1_X(2/n - 1/N_X) - 1_Y(1/N_Y - 2/n)} \\
-	& \leq \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \cdot \left(\norm{1_X} \frac{\abs{n/2 - N_X}}{n/2 N_X} + \norm{1_Y} \frac{\abs{n/2 - N_Y}}{n/2 N_Y}\right) \\
-	& \leq \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \cdot \left(\frac{2\abs{n/2 - N_X}}{\sqrt{N_X} n} + \frac{2\abs{n/2 - N_Y}}{\sqrt{N_Y} n} \right).
-	\end{align*}
-	The reasoning in the proof of Lemma \ref{lem: noise_quadratic_form_on_grid} leads to the bound
-	\begin{equation*}
-	\sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \leq \frac{C n^{1/d}}{\pi},
-	\end{equation*}
-	and a standard application of Hoeffding's inequality to the quantities $N_X$ and $N_Y$ yields, for any $a > 0$
-	\begin{equation*}
-	\abs{N_X - n/2}, \abs{N_Y - n/2} \leq  a \sqrt{n}
-	\end{equation*}
-	with probability at least $1 - 2 \exp(-2a)$. As a result,
-	\begin{equation*}
-	\sup_{\substack{\theta: \norm{D_r \theta}_2 \leq 1, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \cdot \left(\frac{2\abs{n/2 - N_X}}{\sqrt{N_X} n} + \frac{2\abs{n/2 - N_Y}}{\sqrt{N_Y} n} \right) \leq \frac{8a C n^{1/d}}{n\pi}
-	\end{equation*}
-	with probability at least $1 - 2 \exp(-2a)$, which proves the claim.
-\end{proof}
-
-\subsection{Type I error.}
-
-Denote the eigenvalues and eigenvectors of $\Lgrid = \Dgrid^T \Dgrid$ as $\set{\lambda_{k}: k \in [\kappa]^d}$ and $\set{u_{k}: k \in [\kappa]^d}$, respectively.
-
-\begin{lemma}
-	\label{lem: noise_quadratic_form_on_grid}
-	Let $\sigma_1, \ldots, \sigma_n \sim \mathrm{Rademacher}(1/2)$ be independent random variables, and write $\sigma = (\sigma_1, \ldots, \sigma_n)$. Then, 
-	\begin{equation*}
-	\sigma^T \Linvgrid \sigma \leq \frac{(2d)^d\kappa^2}{a\pi^2} \left(8 + 2\pi \log(\sqrt{2} \kappa) \left(\sum_{p = 2}^{d}\kappa^{p}\right) \right)
-	\end{equation*}
-	with probability at least $1  - a$.
-\end{lemma}
-\begin{proof}
-	Taking the eigendecomposition $\Lgrid = U \Lambda U^T$, we can then write $\Linvgrid = U \Lambda^{\dagger} U^T$. Therefore
-	\begin{equation*}
-	\Ebb \bigl(\sigma^T \Linvgrid \sigma\bigr) = \sum_{k \in [\kappa]^d} \frac{\Ebb \bigl(u_{k}^T \sigma\bigr)^2}{\lambda_k}
-	\end{equation*}
-	and as 
-	\begin{equation*}
-	\Ebb \bigl(u_{k}^T \sigma\bigr)^2 = \sum_{i,j = 1}^{n} \Ebb(u_{k_i} u_{k_j} \sigma_i \sigma_j) = \sum_{i = 1}^{n} u_{k_i}^2 = 1
-	\end{equation*}
-	we are left with
-	\begin{equation}
-	\label{eqn: noise_quadratic_form_on_grid_1}
-	\Ebb \bigl(\sigma^T \Linvgrid \sigma\bigr) = \sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \frac{1}{\lambda_k}.
-	\end{equation}
-
-	It is well known \textcolor{red}{HR, CB}, that the $d$-dimensional grid can be written as the Kronecker product of $d$ path graphs, and exploiting this fact as in \textcolor{red}{SWT} we obtain
-	\begin{align*}
-	\lambda_{k} = 4 \sin^2\left(\frac{\pi(k_1 - 1)}{2 \kappa}\right) + \cdots + 4 \sin^2\left(\frac{\pi(k_d - 1)}{2 \kappa}\right).
-	\end{align*}
-	The inequality $\sin(x) \geq x/2$ holds for $x \in [0,\pi/2]$, and so we have
-	\begin{align*}
-	\sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \frac{1}{\lambda_k} & \leq \frac{\kappa^2}{\pi^2} \sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d} (k_j - 1)^2\right)^{-1} \\
-	& \leq \frac{(2d)^d\kappa^2}{\pi^2} \left(8 + 2\pi \log(\sqrt{2} \kappa) \left(\sum_{p = 2}^{d}\kappa^{p - 2}\right) \right) \\
-	& = \frac{(2d)^d\kappa^2}{\pi^2} \left(8 + 2\pi \log(\sqrt{2} \kappa) \left(\sum_{p = 2}^{d}\kappa^{p}\right) \right) \tag{Lemma \ref{lem: grid_eigenvalues}}
-	\end{align*}
-	whence the claim follows by Markov's inequality.
-\end{proof}
-
-\begin{lemma}
-	\label{lem: grid_eigenvalues}
-	There exists a universal constant $c_1 > 0$ such that for any $\kappa \geq 1$ and for all integers $d \geq 3$,
-	\begin{equation*}
-	\sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d} (k_j - 1)^2\right)^{-1} \leq (2d)^d \Biggl(8 + 2 \pi \log (\sqrt{2} \kappa) \biggl(\sum_{p = 2}^{d} \kappa^{p - 2}\biggr) \Biggr)
+	(\thetast)^T \Linv \thetast \geq \frac{1}{\lambda_k}\left(\dotp{\thetast}{\thetast} - \dotp{P_k^{\perp} \thetast}{P_k^{\perp}\thetast}\right)^2
 	\end{equation*}
@@
 \begin{proof}
-	We will prove by induction on $d$. As a base case, consider $d  = 2$. Rewrite
+	Note that $\Lambda^{\dagger}$ is a diagonal matrix, with entries $\rho_{1,1} = 0$ and $\rho_{k,k} = \frac{1}{\lambda_k}$ for $k = 1, \ldots, n$. Therefore,
 	\begin{equation*}
-	\sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d} (k_j - 1)^2\right)^{-1} \leq 4 \sum_{k_1 = 1}^{\kappa - 1} \frac{1}{k_1^2}  + \sum_{k_1 =2}^{\kappa - 1} \sum_{k_2 = 2}^{\kappa - 1} (k_1^2 + k_2^2)^{-1}.
+	(\thetast)^T \Linv \thetast = \sum_{k = 1}^{n} \frac{\dotp{\thetast}{u_k}^2}{\lambda_k}.
 	\end{equation*}
-	Then, upper bounding sums by integrals, we obtain
+	Clearly, for any $k = 1, \ldots, n$,
 	\begin{equation*}
-	\sum_{k_1 = 1}^{\kappa - 1} \frac{1}{k_1^2} \leq 1 +  \int_{x = 1}^{\kappa} \frac{1}{x^2} \dif x \leq 2,
+	\sum_{k = 1}^{n} \frac{\dotp{\thetast}{u_k}^2}{\lambda_k} \geq \frac{\dotp{P_k \thetast}{P_k \thetast}^2}{\lambda_k}
 	\end{equation*}
-	and 
+	and as $\dotp{P_k \thetast}{P_k \thetast} + \dotp{P_k^{\perp} \thetast}{P_k^{\perp} \thetast} = \dotp{\thetast}{\thetast}$, we obtain
 	\begin{align*}
-	\sum_{k_1 = 2}^{\kappa - 1} \sum_{k_2 = 2}^{\kappa - 1} (k_1^2 + k_2^2)^{-1} & \leq \int_{x = 1}^{\kappa} \int_{y = 1}^{\kappa} \frac{1}{x^2 + y^2} \dif x \dif y \\
-	& \leq 2 \pi \log(\sqrt{2} \kappa),
+	\frac{\dotp{P_k \thetast}{P_k \thetast}^2}{\lambda_k} & =  \frac{1}{\lambda_k}\left(\dotp{\thetast}{\thetast} - \dotp{P_k^{\perp} \thetast}{P_k^{\perp}\thetast}\right)^2.
 	\end{align*}
-	so that
-	\begin{equation*}
-	\sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d} (k_j - 1)^2\right)^{-1} \leq 8 + 2 \pi \log(\sqrt{2} \kappa).
-	\end{equation*}
-	and the base case is shown.
-
-	We proceed by induction. For $d \geq 3$ we write
-	\begin{align*}
-	\sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d} (k_j - 1)^2\right)^{-1} & \leq 2d \sum_{\substack{k \in [\kappa]^{d - 1}, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d - 1} (k_j - 1)^2\right)^{-1} + \sum_{k_1 = 2}^{\kappa - 1} \cdots \sum_{k_d = 2}^{\kappa - 1} \left(\sum_{j = 1}^{d} k_j^2\right)^{-1} \\
-	& \leq 2d \left( (2d)^{d - 1}\Biggl(8 + 2 \pi \log \sqrt{2} \kappa \biggl(\sum_{p = 2}^{d - 1} \kappa^{p - 2}\biggr) \Biggr)\right) + \sum_{k_1 = 2}^{\kappa - 1} \cdots \sum_{k_d = 2}^{\kappa - 1} \left(\sum_{j = 1}^{d} k_j^2\right)^{-1} \\
-	& = \left( (2d)^{d}\Biggl(8 + 2 \pi \log \sqrt{2} \kappa \biggl(\sum_{p = 2}^{d - 1} \kappa^{p - 2}\biggr) \Biggr)\right) + \sum_{k_1 = 2}^{\kappa - 1} \cdots \sum_{k_d = 2}^{\kappa - 1} \left(\sum_{j = 1}^{d} k_j^2\right)^{-1}
-	\end{align*}
-	and again upper bounding sums by integrals, we obtain
-	\begin{align*}
-	\sum_{k_1 = 2}^{\kappa - 1} \cdots \sum_{k_d = 2}^{\kappa - 1} \left(\sum_{j = 1}^{d} (k_j)^2\right)^{-1} & \leq \int_{1}^{\kappa} \cdots \int_{1}^{\kappa} \frac{1}{x_1^2 + \cdots + x_d^2} \dif x_d \ldots \dif x_1 \\
-	& \leq \kappa^{d-2} \int_{1}^{\kappa} \int_{1}^{\kappa} \frac{1}{x^2 + y^2} \dif x \dif y \\
-	& \leq \kappa^{d-2} 2 \pi \log(\sqrt{2} \kappa)
-	\end{align*}
-	and the proof is complete.
-\end{proof}
-
-\paragraph{Type II error.}
-
-Lemma \ref{lem: approx_error_lb_1} will be useful in lower bounding the \textcolor{red}{approximation error} of the $T_{LS}$ test statistic.
-Write $P_1^{\perp}$ for the projection operator onto the subspace of $\Reals^n$ orthogonal to the constant vector $\mathbf{1}$.
-
-\begin{lemma}
-	\label{lem: approx_error_lb_1}
-	For any vector $v \in \Reals^n$ and Laplacian matrix $L$ of a connected graph $G$,
-	\begin{equation*}
-	v^T \Linv v \geq \frac{(v^T P_{1}^{\perp}v)^2}{v^T L v}
-	\end{equation*}
-\end{lemma}
-\begin{proof}
-	We can expand $v^T P_1^{\perp} v  = v^T (L)^{1/2}(\Linv)^{1/2} v$, for any matrix square root of $L$ and $\Linv$. The statement follows by Cauchy-Schwarz.
 \end{proof}

### 2019-06-07 notes/weekly_notes/6.5.19.NotesForWeek.tex
weekly notes updated

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbb{I}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 6/1/19 - 6/5/19}
\author{Alden Green}
\date{\today}
\maketitle

Let $f,g$ be $L^2$ density functions supported on the $d$-dimensional unit cube $[0,1]^d$, meaning
\begin{equation*}
\int_{[0,1]^d} f(x)^2 dx, \int_{[0,1]^d} g(x)^2 dx < \infty, \quad \text{and} \quad \int_{[0,1]^d} f(x) dx, \int_{[0,1]^d} g(x) dx = 1.
\end{equation*}
with $f(x), g(x) > 0$ for all $x \in [0,1]^d$.

We observe data $(X,\ell)$, a design matrix and associated labels, specified as follows. We consider a fixed design setting, and more precisely we fix $X$ to be a grid over the unit cube. Formally, for a given sample size $n \in \mathbb{N}$ and $\kappa = n^{1/d}$, let $X = \set{x_k: k \in [\kappa]^d}$ with $x_k = k/\kappa$. Our labels are then $\ell = \set{\ell_k: k \in [\kappa]^d}$, with $\ell_k$ independent random variables sampled according to
\begin{equation*}
\ell_k = 
\begin{cases}
1, ~ & \textrm{with probability}~ \frac{f(x_k)}{f(x_k) + g(x_k)}, \\
-1 ~ & \textrm{with probability}~ \frac{g(x_k)}{f(x_k) + g(x_k)}
\end{cases}
\end{equation*}
where dividing by $f(x) + g(x)$ is permissible as $f(x), g(x) > 0$ for all $x \in [0,1]^d$.

Our statistical goal is hypothesis testing: that is, we wish to construct a test function $\phi$ which differentiates between
\begin{equation*}
\mathbb{H}_0: f = g \text{ and } \mathbb{H}_1: f \neq g.
\end{equation*}

For a given function class $\Hclass$, some $\epsilon > 0$, and test function $\phi$ a Borel measurable function of the data with range $\set{0,1}$, we evaluate the quality of the test using \emph{worst-case risk}
\begin{equation*}
R_{\epsilon}^{(t)}(\phi; \Hclass) = \sup_{f \in \Hclass} \Ebb_{f,f}^{(t)}(\phi) + \sup_{ \substack{f,g \in \Hclass \\ \delta(f,g) \geq \epsilon } } \Ebb_{f,g}^{(t)}(1 - \phi)
\end{equation*} 
where 
\begin{equation*}
\delta^2(f,g) = \int_{\D} (f - g)^2 dx.
\end{equation*}
We will consider those density functions which belong to $\mathcal{H} := \mathcal{H}_{1}^d(L)$, the $d$-dimensional Lipschitz ball with norm $L$. Formally, a function $f \in \mathcal{H}$ if for all $x,y \in [0,1]^d$, $\abs{f(x) - f(y)} \leq L \norm{x - y}$.\footnote{Note that the Lipschitz requirement, along with the restriction that $f$ be a density function over a bounded domain, imply that $f$ has finite $L^2$ norm.}

\section{Discretization error in fixed design setting.}
Our first result is a lower bound. We exhibit density functions $f,g \in \mathcal{H}_1^{d}(\pi/2)$ such that $\delta^2(f,g) = n^{-2/d}$ for which, assuming fixed design $X$ as we have done, no test can distinguish $f - g$ from the $0$ function. For simplicity, assume $\kappa$ is an odd integer.
\begin{proposition}
	\label{prop: discretization_error}
	For $x \in [0,1]^d$, let $f,g \in \mathcal{H}_1^d(\pi/2)$ be density functions given by
	\begin{equation*}
	f(x) = 1 + \frac{1}{2\kappa}\prod_{i = 1}^{d} \sin(\kappa \pi x_i), \quad g(x) = 1 - \frac{1}{2\kappa}\prod_{i = 1}^{d} \sin(\kappa \pi x_i)
	\end{equation*}
	Then, $\delta^2(f,g) = n^{-2/d}$, and $f(x_k) = g(x_k)$ for every $k \in [\kappa]^d$.
\end{proposition}
As a result, $(f - g)(x_k) = 0$ for every $x_k \in X$, and so the labels $\ell$ have the same distribution as they would under the null hypothesis $f = g$. 
\begin{proof}[Proof of Proposition \ref{prop: discretization_error}]
	We begin by showing that $f$ and $g$ are density functions with Lipschitz norm bounded by $1$. We compute the integral of $f$ over its support,
	\begin{align*}
	\int_{[0,1]^d} f(x) \dx & = 1 + \frac{1}{2\kappa} \int_{[0,1]^d} \prod_{i = 1}^{d} \sin(\kappa \pi x_i) \dx_i \\
	& = 1 + \frac{1}{2\kappa} \left(\int_{0}^{1} \sin(\kappa \pi x) \dx\right)^d \\
	& = 1 + \frac{1}{2 \kappa} \left(-\cos(\kappa \pi x) \Bigr|_{0}^{1}\right) = 1.
	\end{align*}
	where the last equality follows as $\kappa$ is an odd integer. Similar calculations follow for $g$, and therefore $f$ and $g$ are density functions.

	Next, we upper bound the Lipschitz norm of $f$ (again, similar calculations will hold for $g$). Taking the partial derivative of $f$ along any coordinate $x_i$, we obtain
	\begin{align*}
	\abs{\frac{\partial f}{\partial x_i}} & = \abs{\frac{\kappa \pi}{2 \kappa} \cos(\pi \kappa x_i) \prod_{j \neq i}^{d} \sin(\kappa \pi x_j)} \\
	& \leq \frac{\pi}{2}.
	\end{align*}
	Since this holds for each $i = 1, \ldots, d$, $f$ has Lipschitz norm upper bounded by $\pi/2$.

	Now, we show $\delta^2(f - g) \geq 2^{-(d + 2)} n^{-2/d}$,
	\begin{align*}
	\int_{[0,1]^d} (f(x) - g(x))^2 \dx & = \frac{1}{4 \kappa^2} \int_{[0,1]^d} \left(\prod_{i = 1}^{d} \sin(\kappa \pi x_i)\right)^2 \dx_i \\
	& = \frac{1}{4\kappa^2} \left( \int_{0}^{1} \sin^2(\kappa \pi x) \dx \right)^d \\
	& = \frac{1}{4\kappa^2} \left( \int_{0}^{1} \frac{1 - \cos(2 \kappa \pi x)}{2} \dx \right)^d \\ 
	& = \frac{1}{4\kappa^2} \left(\frac{1 - \frac{\sin(2 \kappa \pi x)}{2 \kappa \pi}}{2} \biggr|_{0}^{1} \right)^d \\
	& = \frac{1}{4 2^d \kappa^2}
	\end{align*}
	with the last equality following from the fact $\kappa$ is an integer.

	Finally, for any $k \in [\kappa]^d$ and $x_k = k/\kappa$, we have
	\begin{equation*}
	\sin(\kappa \pi (x_k)_i) = \sin(\pi k_i) = 0
	\end{equation*}
	and so $f(x_k) = 1 = g(x_k)$. 
\end{proof}

\begin{remark}
	We note that by \citep{ariascastro2018}, the \textcolor{red}{minimax rate} for hypothesis testing over the Holder ball $\mathcal{H}_{\beta}^d(L)$ is of order $n^{-4\beta/(4\beta + d)}$. In particular, the aforementioned Lipschitz ball is simply $\mathcal{H}_1^d(L)$, and the minimax rate becomes $n^{-4/(4 + d)}$. We note that this is faster than the lower bound $n^{-2/d}$ provided by Proposition \ref{prop: discretization_error}. The correct interpretation is that the discretization error due to considering fixed grid design (and choosing negative examples with knowledge of that design) dominates other forms of error (i.e. squared bias and variance). Interestingly, this appears not to be true for estimation in the analogous case. Here, the minimax rate of estimation over the Lipschitz ball with $L^2$ risk is $n^{-2/(2 + d)}$, which dominates the exhibited discretization error $n^{-2/d}$. 
\end{remark}

\section{Sub-optimality of the discrete Sobolev IPM for fixed design.}

For a function $h: [0,1]^d \to \Reals$, write $\theta^{h} := (h(x_k): x_k \in X)$ for the evaluation of $h$ at the data. In the previous section, we exhibited a function $h := f - g$ such that $||h||_{L^2} \soom n^{-1/d}$, but $\norm{\theta^{h}}_2 = 0$. To avoid this issue, let us shift slightly our requirements on $f,g$, to be placed directly on discrete norms rather than continuous norms. In particular, let $D$ be the incidence matrix of a lattice graph $G = (V,E)$ over the design points $X$, where $V = [\kappa]^d$ and $(k,k') \in E$ if $\norm{k - k'}_1 = 1$. Now, we will consider those density functions $f: [0,1]^d \to \Reals$ such that the discrete Sobolev norm is bounded $\norm{D\theta^f}_2 \leq n^{1/2 - 1/d}$. In particular, letting
\begin{equation*}
\overline{\mathcal{W}}_{1}^d(L) = \set{f: \norm{D \theta_f}_2 \leq n^{1/2 - 1/d}}
\end{equation*}
for $f, g \in \overline{\mathcal{W}}_{1}^d(L)$ we now wish to test whether $\norm{\theta^f - \theta^g}_2 = 0$.

A potential test statistic in this context is the discrete Sobolev integral probability metric (IPM), formally defined as
\begin{equation*}
T_{S} := \sup_{\theta: ||D\theta||_2 \leq C, \theta^T \mathbf{1} = 0} 
\frac{1}{n}\dotp{\theta}{\ell}
\end{equation*}
Letting $L = D^T D$ be the discrete Laplacian over the lattice, and $\Linv$ be the pseudoinverse of $L$, a quick calculation yields $T_S = \frac{C}{n} \sqrt{\ell^T \Linv \ell}$. Is $T_{S}$ an optimal test over vectors $\theta^f, \theta^g$ in the discrete Sobolev ball, with respect to the empirical norm $\norm{\theta^f - \theta^g}_2$? Propositions \ref{prop: suboptimal_IPM}, \ref{prop: type_I_error} and \ref{prop: k_eigenvector_test} show that it is not. Hereafter, we let $\sqrt{\kappa}$ be an integer.
\begin{proposition}
	\label{prop: suboptimal_IPM}
	Let $f,g \in \overline{\mathcal{W}}_1^d(\pi/2)$ be functions given by
	\begin{equation*}
	f(x) = 1 + \frac{1}{2\sqrt{\kappa}}\prod_{i = 1}^{d} \cos\left(x_i \sqrt{\kappa} \pi - \frac{\pi}{2 \sqrt{\kappa}}\right), \quad g(x) =  1 - \frac{1}{2\sqrt{\kappa}}\prod_{i = 1}^{d} \cos\left(x_i \sqrt{\kappa} \pi - \frac{\pi}{2 \sqrt{\kappa}}\right)
	\end{equation*}
	Then the following statements hold:
	\begin{itemize}
		\item $\norm{\theta^f - \theta^g}_2 = 2^{-(d+2)/2} n^{1/2 - 1/2d}$.
		\item $\norm{D \theta^f}_2 \leq (2^{-(d+2)/2}\sqrt{d}\pi) n^{1/2 - 1/d}$.
		\item $\sup_{\theta: ||D\theta||_2 \leq C, \theta^T \mathbf{1} = 0} 
		\frac{1}{n}\dotp{\theta}{\theta^f - \theta^g} \leq C(d^{-1/2} 2^{-d/2}) n^{-1/2}$
	\end{itemize}
\end{proposition}
\begin{proof}
	Write $\theta^h : = \theta^f - \theta^g$. In the one-dimensional case (i.e $G$ being the path graph), it is well known that for any $k \in 1,\ldots, \kappa$, the vector 
	\begin{equation*}
	v = (v_k)_{j=1}^{\kappa}, v_k = \cos\left(\frac{(j - 1/2)k\pi}{\kappa}\right)
	\end{equation*}
	is a (scaled) eigenvector of $L$ with eigenvalue $4 \sin^2(k \pi/2\kappa)$; precisely
	\begin{equation*}
	L v = (4 \sin^2(j \pi/2 \kappa)) v,~ \text{and} \norm{v} = (\kappa/2)^{1/2}
	\end{equation*}
	As $\sqrt{\kappa}$ is an integer, one can use standard facts about Cartesian product graphs to derive that $\theta^h$ is itself a (scaled) eigenvector of $L$; precisely
	\begin{equation*}
	L \theta^h = (4d \sin^2(\pi \sqrt{\kappa}/2{\kappa})) \theta^h,~ \textrm{and} \norm{\theta^h}_2 = \frac{1}{2 \sqrt{\kappa}}\left(\frac{\kappa}{2}\right)^{d/2}.
	\end{equation*}
	The second equality proves the first claim. Both equalities together yield
	\begin{align*}
	\norm{D \theta^h}_2 & = 2 \sqrt{d} \sin(\pi \sqrt{\kappa}/2\kappa) \frac{1}{2 \sqrt{\kappa}} \left(\frac{\kappa}{2}\right)^{d/2} \\
	& \leq \frac{1}{2 \kappa}\sqrt{d} \pi \left(\frac{\kappa}{2}\right)^{d/2}
	\end{align*}
	which proves the second claim. Finally, as we have noted $\sup_{\theta: ||D\theta||_2 \leq C, \theta^T \mathbf{1} = 0} 
	\frac{1}{n}\dotp{\theta}{\theta^f - \theta^g} = \frac{C}{n} \sqrt{(\theta^h)^T \Linv \theta^h}$, and since
	\begin{align*}
	\sqrt{(\theta^h)^T \Linv \theta^h} & = \frac{\norm{\theta^h}_2}{2 \sqrt{d} \sin(\pi \sqrt{\kappa}/2{\kappa})} \\
	& \leq \frac{1}{\sqrt{d}}\left(\frac{\kappa}{2}\right)^{d/2}
	\end{align*}
	the third claim follows.
\end{proof}

Proposition \ref{prop: type_I_error} demonstrates that under the null hypothesis, $T_S$ scales at approximately the same rate, $n^{-1/2}$, as it does for the choice of $f$ and $g$ in Proposition \ref{prop: suboptimal_IPM}.
\begin{proposition}
	\label{prop: type_I_error}
	Fix $0 < \delta < 1$. If $\theta^f - \theta^g = 0$, there exists a constant $C(d,\delta)$ based only on $d$ and $\delta$, such that with probability at least $1 - \delta$,
	\begin{equation*}
	T_{S} \geq C(d,\delta) \frac{C}{\sqrt{n}}.
	\end{equation*}
\end{proposition}

To prove Proposition \ref{prop: type_I_error}, it will be helpful to introduce multiindex notation to refer to the eigenvalues and eigenvectors of $L$. Specifically, for $k,j \in [\kappa]^d, k = (k_1, \ldots, k_d), j = (j_1, \ldots, j_d)$, it is well known that the eigenvector $v_k$ has coordinates $(v_k)_j$ and corresponding eigenvalue $\lambda_k$ given by
\begin{align*}
(v_k)_j & = \left(\frac{2}{\kappa}\right)^{d/2} \prod_{i = 1}^{d} \cos\left(\frac{(j_i - 1/2) k_i \pi}{\kappa}\right) \\
\lambda_k & = 4 \sum_{i = 1}^{d} \sin^2\left(\frac{\pi(k_i - 1)}{2\kappa}\right).
\end{align*}
\begin{proof}
	Throughout, let $c$ be a constant which depends only on $d$ and $\delta$, which may change from line to line. We will show that under the null hypothesis $\theta^f - \theta^g = 0$, $\Ebb(\ell^T \Linv \ell) = c n$, in which case the result follows from (e.g.) an application of Markov's inequality.

	Under the null hypothesis $\theta^f - \theta^g = 0$, we note that $\ell$ consists of $n$ i.i.d Rademacher random variables. Hence
	\begin{align*}
	\Ebb(\ell^T \Linv \ell) & = \sum_{k \in [\kappa]^d} \frac{(\ell^T v_k)^2}{\lambda_k} \\
	& = \sum_{k \in [\kappa]^d} \frac{1}{\lambda_k} \\
	& = \frac{1}{4} \sum_{k \in [\kappa]^d} \frac{1}{\sum_{i = 1}^{d} \sin^2\left(\frac{\pi(k_i - 1)}{2\kappa}\right)} \\
	& \geq \frac{\kappa^2}{\pi^2} \sum_{k \in [\kappa]^d} \frac{1}{\sum_{i = 1}^{d} (k_i - 1)^2} \\
	& \geq \frac{\kappa^2}{\pi^2} \int_{x \in [1,\kappa]^d} \frac{1}{\sum_{i = 1}^d x_i^2} \dx \\
	& = c \kappa^2 \int_{0}^{\kappa} \frac{1}{r^2} r^{d - 1} \dr \\
	& = c \kappa^2 \kappa^{d - 2} = c \kappa^d. 
	\end{align*}
	and the claim follows since $\kappa^d = n$.
\end{proof}

Together, Propositions \ref{prop: suboptimal_IPM} and \ref{prop: type_I_error} show that there exists functions $f,g \in \overline{\mathcal{W}}_1^d$ with empirical norm $\norm{\theta^f - \theta^g}_2 \gtrsim n^{1/2 - 1/2d}$ such that $T_S$ cannot distinguish $\theta^f - \theta^g$ from the zero vector.

Now, we turn our attention to a different test. Denote by $V_m = [v_1 \ldots v_k]$ the $n \times m$ matrix containing the first $m$ eigenvectors of $L$ (that is, the eigenvectors corresponding to the smallest $m$ eigenvalues of $L$, excluding the eigenvalue $\lambda = 0$). The \emph{Laplacian eigenmaps} estimate of $\Ebb(\ell) = \frac{\theta^f - \theta^g}{\theta^f + \theta^g}$ is the projection of $\ell$ onto the first $m$ eigenvectors of $L$, 
\begin{equation*}
\widehat{\theta}^{LE} = (V_m V_m^T) y.
\end{equation*}

Proposition \ref{prop: k_eigenvector_test} shows that a test based on $\norm{\widehat{\theta}^{LE}}_2$ can distinguish any two $f,g \in \overline{\mathcal{W}}_1^d$ such that $\norm{\theta^f - \theta^g}_2$ is at least roughly $n^{1/2 - 1/(d + 2)}$. It is a consequence of the estimation results derived in \cite{sadhanala2016}.

\begin{proposition}
	\label{prop: k_eigenvector_test}
	Fix $0 < \delta < 1$, and write $\widetilde{\theta}^h := (\theta^f - \theta^g)/(\theta^f - \theta^g)$. Assume $f,g \in \overline{\mathcal{W}}_1^d(L)$ satisfy some lower bound $f(x), g(x) \geq p_{\min}$ for all $x \in [0,1]^d$. Then, there exists a constant $C(L, d,\delta,p_{\min})$ such that
	\begin{itemize}
		\item $\norm{\widehat{\theta}^{LE}}_2 \geq \norm{\widetilde{\theta}^h}_2 -  C(L, d,\delta, p_{\min})n^{1/2 -1/(d + 2)}$
		\item $\norm{\widehat{\theta}^{LE}}_2 \leq \norm{\widetilde{\theta}^h}_2 +  C(L, d,\delta, p_{\min})n^{1/2 -1/(d + 2)}$
	\end{itemize}
	each with probability at least $1 - \delta$.
\end{proposition}
\begin{proof}
	We begin by noting that
	\begin{equation*}
	\norm{D \widetilde{\theta}^h}_2 \leq \frac{L}{p_{\min}}
	\end{equation*}
	In \cite{sadhanala2016}, it is shown that for $y = \widetilde{\theta}^h + w$, where $w$ is i.i.d Normal noise with unit variance,
	\begin{equation*}
	\norm{\widehat{\theta}^{LE} - \widetilde{\theta}^h}_2^2 \leq c n \frac{1}{n^{\frac{2}{d+2}}}
	\end{equation*}
	where $c$ is a constant depending on $d$, $L$, and $p_{\min}$. We take for granted that this result can be extended to the case $\ell = \widetilde{\theta}^h + w$, where $w$ consists of independent but not necessarily identically distributed noise with mean zero and variance no greater than $1$. Both claims then follow from the triangle inequality.
\end{proof}

\clearpage
\bibliographystyle{plain}
\bibliography{../../graph_testing_bibliography}
\end{document}

### 2019-06-11 notes/weekly_notes/6.9.19.NotesForWeek.tex
added weekly notes

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbb{I}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 6/9/19 - 6/11/19}
\author{Alden Green}
\date{\today}
\maketitle

Let $f,g$ be $L^2$ density functions supported on the $d$-dimensional unit cube $[0,1]^d$, meaning
\begin{equation*}
\int_{[0,1]^d} f(x)^2 dx, \int_{[0,1]^d} g(x)^2 dx < \infty, \quad \text{and} \quad \int_{[0,1]^d} f(x) dx, \int_{[0,1]^d} g(x) dx = 1.
\end{equation*}
with $f$,$g$ bounded away from $0$, $f(x), g(x) > p_{\min}$ for all $x \in [0,1]^d$.

We observe data $(X,\ell)$, a design matrix and associated labels, specified as follows. We let $x_1, \ldots, x_n$ be the rows of $X$, each sampled independently from $\mu = (f + g)/2$. For each $i = 1,\ldots,n$, we then sample $\ell_i$ according to
\begin{equation*}
\ell_i =
\begin{cases}
1,~ \textrm{with probability} \frac{f(x_i)}{f(x_i) + g(x_i)} \\
-1,~ \textrm{with probability} \frac{g(x_i)}{f(x_i) + g(x_i)}
\end{cases}
\end{equation*}
and let $\ell = (\ell_i)$.

Our statistical goal is hypothesis testing: that is, we wish to construct a test function $\phi$ which differentiates between
\begin{equation*}
\mathbb{H}_0: f = g \text{ and } \mathbb{H}_1: f \neq g.
\end{equation*}

For a given function class $\Hclass$, some $\epsilon > 0$, and test function $\phi$ a Borel measurable function of the data with range $\set{0,1}$, we evaluate the quality of the test using \emph{worst-case risk}
\begin{equation*}
R_{\epsilon}^{(t)}(\phi; \Hclass) = \sup_{f \in \Hclass} \Ebb_{f,f}^{(t)}(\phi) + \sup_{ \substack{f,g \in \Hclass \\ \delta(f,g) \geq \epsilon } } \Ebb_{f,g}^{(t)}(1 - \phi)
\end{equation*} 
where 
\begin{equation*}
\delta^2(f,g) = \int_{\D} (f - g)^2 dx.
\end{equation*}
We will consider those density functions which belong to $\mathcal{H} := \mathcal{H}_{1}^d(L)$, the $d$-dimensional Lipschitz ball with norm $L$. Formally, a function $f \in \mathcal{H}$ if for all $x,y \in [0,1]^d$, $\abs{f(x) - f(y)} \leq L \norm{x - y}$.\footnote{Note that the Lipschitz requirement, along with the restriction that $f$ be a density function over a bounded domain, imply that $f$ has finite $L^2$ norm.}

\section{Sieve-based test.}

For a radius $r > 0$ to be determined later, the $r$-neighborhood graph $G = (V,E)$ consists of vertices $V = [n]$ corresponding to the $n$ data points, and edges $E = \set{(u,v): \norm{x_u - x_v} \leq r}$. Write $D$ for the incidence matrix of this graph, and $L = D^T D$ for the graph Laplacian. Denote the singular value decomposition of $D$ by $D = U \Lambda^{1/2} V^T$, where $U$ and $V$ are orthonormal matrices and $\Lambda$ is a diagonal matrix with entries $0 = \lambda_1 \leq \lambda_2 \leq \lambda_3 \leq \ldots \leq \lambda_n$, so that $L = V \Lambda V^T$. For $k = 1,\ldots,n$, write $V_k = (v_1 \ldots v_k)$ for the $n \times k$ matrix containing the first $k$ columns of $V$--that is, the first $k$ eigenvectors of $L$. Our first test statistic will be as follows:

\begin{equation}
\label{eqn: laplacian_eigenmaps_statistic}
T_{LE}^{(k)} := \sup_{\substack{\theta \in \mathrm{col}(V_k), \\ \norm{\theta} = 1} } \dotp{\theta}{\ell}^2 = \ell^T V_k V_k^T \ell.
\end{equation}

\subsection{Assumptions}
The following are statements I wish to hold in order for the later theory to follow, but I cannot yet prove:
\begin{enumerate}
	\item There exists a constant $c_1 > 0$ -- which may depend on dimension $d$ but not on $n$ --  such that for all $h \in 1, \ldots,k$ and all $i = 1,\ldots,n$, 
	\begin{equation}
	\label{eqn: asmp_1}
	\abs{v_{hi}} \leq \frac{c}{\sqrt{n}}
	\end{equation}
\end{enumerate}

If $v_k$ were eigenvectors of the grid--rather than of the neighborhood graph--the first item would hold with $c = 2^{d/2}$.

\subsection{Main result.}
Let $d = 3,4$, choose $r = c(\log(n)/n)^{1/d}$ to satisfy the condition of Lemma \ref{lem: approximation_error}, and let $k = n^{2d/(d + 4)}$. Let 
\begin{equation*}
\phi_{LE} =
\begin{cases}
1,~ & \text{if}~ T_{LE}^{(k)} \geq k + a\sqrt{2k + c^4\left(k + \frac{k^2}{n}\right)} \\
0,~ & \text{otherwise}.
\end{cases}
\end{equation*}
The following theorem holds.
\begin{theorem}
	There exists some number $c > 0$ constant in sample size $n$ such that for any $\epsilon \geq c n^{-2/(4 + d)}$, we have
	\begin{equation*}
	R_{\epsilon}^{(t)}(\phi_{LE}; \Hclass_1^{d}(L)) \leq \frac{1}{a^2}.
	\end{equation*}
	when $d = 3$ or $4$. 
\end{theorem}
\begin{proof}

\end{proof}

\subsection{Type I error}

Under the null hypothesis, $\ell$ consists of $n$ independent and identically distributed Rademacher random variables. Using this fact, we can bound the first two moments of $T_{LE}$ under the null hypothesis.

\begin{lemma}
	If $f = g$, then
	\begin{equation}
	\Ebb(T_{LE}^{(k)}) = k
	\end{equation}
	and
	\begin{equation}
	\label{eqn: variance_noise_typeI}
	\Var(T_{LE}^{(k)}) \leq 2k + c_1^4\left(k + \frac{2k^2}{n}\right)
	\end{equation}
\end{lemma}
\begin{proof}
	Taking the expectation first, we have
	\begin{align*}
	\Ebb(T_{LE}^{(k)}) & = \sum_{j = 1}^{k} \Ebb((\ell^T v_j)^2) \\
	& = \sum_{j = 1}^{k} \sum_{i = 1}^{n} \sum_{i' = 1}^{n} \Ebb(\ell_i \ell_{i'}) v_{ji} v_{ji'} \\
	& =  \sum_{j = 1}^{k} \sum_{i = 1}^{n} v_{ji}^2 = k.
	\end{align*}

	We defer the proof of \eqref{eqn: variance_noise_typeI} until the following subsection, where we show it in greater generality.
\end{proof}

\subsection{Type II error.}

Write $\ell =: \thetast + w$, where $\thetast = (\thetast_i)$ is defined by
\begin{equation*}
\thetast_i = \frac{f(x_i) - g(x_i)}{f(x_i) + g(x_i)}
\end{equation*}
and $w = (w_i)$ therefore consists of $n$ independent (although not identically distributed) zero-mean noise terms. We expand
\begin{equation}
\label{eqn: LE_expansion}
\ell^T V_k V_k^T \ell = (\thetast)^T V_k V_k^T \thetast + w^T V_k V_k w + 2 w^T V_k V_k^T \thetast,
\end{equation}
and focus at first our attention on the second term in \eqref{eqn: LE_expansion}.

\begin{lemma}
	Assuming \eqref{eqn: asmp_1} holds,
	\begin{equation}
	\label{eqn: expectation_noise_typeII}
	\Ebb \left( w^T V_k V_k w\right) \geq k - \frac{kc_1^2 \delta^2(f,g)}{p_{\min}^2}
	\end{equation}
	and
	\begin{equation}
	\label{eqn: variance_noise}
	\Var(T_{LE}^{(k)}) \leq 2k + c_1^4\left(k + \frac{2k^2}{n}\right)
	\end{equation}
\end{lemma}
\begin{proof}
	We note that, conditional on $X$,
	\begin{equation*}
	w_i = 
	\begin{cases}
	2\frac{g(x_i)}{f(x_i) + g(x_i)},~ & \text{with probability}~ \frac{f(x_i)}{f(x_i) + g(x_i)} \\
	-2\frac{f(x_i)}{f(x_i) + g(x_i)},~ & \text{with probability}~ \frac{g(x_i)}{f(x_i) + g(x_i)}
	\end{cases}
	\end{equation*}

	We first show \eqref{eqn: expectation_noise_typeII}. Expanding the quadratic form as a double sum, and using the law of iterated expectation, we have
	\begin{align}
	\label{eqn: expectation_noise_typeII_1}
	\Ebb \left( w^T V_k V_k w\right) & = \sum_{h = 1}^{k} \sum_{i = 1}^{n} \sum_{j = 1}^n \Ebb \bigl(w_i w_j v_{hi} v_{hj} \bigr) \nonumber \\
	& = \sum_{h = 1}^{k} \sum_{i = 1}^{n} \sum_{j = 1}^n \Ebb \biggl( \Ebb\Bigl(w_i w_j \bigl\vert x_i, x_j\Bigr) v_{hi} v_{hj}\biggr) \nonumber \\
	& = \sum_{h = 1}^{k} \sum_{i = 1}^{n} \Ebb \biggl( \Ebb\Bigl(w_i^2 \bigl\vert x_i\Bigr) v_{hi}^2\biggr).
	\end{align}
	The conditional expectation $\Ebb\Bigl(w_i^2 \bigl\vert x_i\Bigr)$ can be directly computed,
	\begin{align*}
	\Ebb\Bigl(w_i^2 \bigl\vert x_i\Bigr) & = 4\left(\frac{g(x_i)}{f(x_i) + g(x_i)}\right)^2 \frac{f(x_i)}{f(x_i) + g(x_i)} + 4\left(\frac{f(x_i)}{f(x_i) + g(x_i)}\right)^2 \frac{g(x_i)}{f(x_i) + g(x_i)} \\
	& = 4 \frac{f(x_i)g(x_i)(f(x_i) + g(x_i))}{(f(x_i) + g(x_i))^3} \\
	& = 4 \frac{f(x_i)g(x_i)}{(f(x_i) + g(x_i))^2}.
	\end{align*}
	and plugging this back into \eqref{eqn: expectation_noise_typeII_1}, we obtain
	\begin{align*}
	\Ebb \left( w^T V_k V_k w\right) & = \sum_{h = 1}^{k} \sum_{i = 1}^{n} \Ebb \left(4 \frac{f(x_i) g(x_i)}{(f(x_i) + g(x_i))^2} v_{hi}^2\right) \\
	& = \sum_{h = 1}^{k} \sum_{i = 1}^{n} \Ebb \left(v_{hi}^2 \left(1 - \left(\frac{f(x_i) - g(x_i)}{f(x_i) + g(x_i)} \right)^2 \right)\right) \\
	& \geq k - \frac{c^2}{n}\sum_{h = 1}^{k} \sum_{i = 1}^{n} \Ebb \left(\frac{f(x_i) - g(x_i)}{f(x_i) + g(x_i)} \right)^2 \\
	& \geq k - \frac{kc_1^2 \delta^2(f,g)}{p_{\min}^2}
	\end{align*}
	and \eqref{eqn: expectation_noise_typeII} is shown.

	We now show \eqref{eqn: variance_noise}. It will be helpful to note that for all $i,i',j,j' \in [n]$,
	\begin{equation*}
	\mathrm{Cov}(w_iw_j, w_{i'}w_{j'} | X) = 
	\begin{cases}
	0, & \text{if}~ \set{i,i',j,j'}~ \textrm{has four distinct elements} \\
	0, & \text{if}~ \set{i,i',j,j'}~ \textrm{has three distinct elements} \\
	0, & \text{if}~ \set{i,i',j,j'}~ \textrm{has two distinct elements and}~ i = j \\
	\Ebb(w_i^2|X) \Ebb(w_j^2|X), & \text{if}~ \set{i,i',j,j'}~ \textrm{has two distinct elements and}~ i \neq j \\
	\Var(w_i^4|X), & \text{if}~ \set{i,i',j,j'}~ \textrm{has one distinct element.}
	\end{cases}
	\end{equation*}
	and additionally
	\begin{equation*}
	\Ebb(w_iw_j|X) =
	\begin{cases}
	0, & \text{if}~ i \neq j \\
	E(w_i^2|X), & \text{if}~ i = j.
	\end{cases}
	\end{equation*}
	Then, by the law of total covariance,
	\begin{align*}
	\mathrm{Var}(w^T V_k V_k^T w) & = \sum_{h,h' = 1}^{k} \sum_{i,i' = 1}^{n} \sum_{j,j' = 1}^{n} \mathrm{Cov} \bigl(w_iw_jv_{hi} v_{hj}, w_{i'}w_{j'}v_{h'i'} v_{h'j'}\bigr) \\
	& = \sum_{h,h' = 1}^{k} \sum_{i,i' = 1}^{n} \sum_{j,j' = 1}^{n} \Ebb \Bigl(  v_{hi} v_{hj} v_{h'i'} v_{h'j'} \mathrm{Cov} \bigl(w_iw_j, w_{i'}w_{j'} |X\bigr)\Bigr) + \\
	& \quad \quad \mathrm{Cov} \Bigl( v_{hi}v_{hj} \Ebb(w_iw_j|X), v_{h'i'}v_{h'j'} \Ebb(w_i'w_j'|X)  \Bigr) \\
	& = 2 \sum_{h,h' = 1}^{k} \sum_{i \neq j}^{n} \Ebb\Bigl( v_{hi} v_{hj} v_{h'i} v_{h'j} \Ebb(w_i^2|X) \Ebb(w_j^2|X) \Bigr) + \\
	& \quad \quad \sum_{h,h' = 1}^{k} \sum_{i = 1}^{n} \Ebb \left(v_{hi}^2 v_{h'i}^2 \Var(w_i^4|X)\right) +  \\
	& \quad \quad \quad \sum_{h,h' = 1}^{k} \sum_{i,i' = 1}^{n} \mathrm{Cov} \bigl( v_{hi} ^2 \Ebb(w_i^2|X), v_{h'i'}^2 \Ebb(w_{i'}^2|X)\bigr)
	\end{align*}
	Then, using the upper bound $\Ebb(w_i^2|X),\Ebb(w_j^2|X) \leq 1$, we can bound the first term
	\begin{align*}
	\sum_{h,h' = 1}^{k} \sum_{i \neq j}^{n} \Ebb\Bigl( v_{hi} v_{hj} v_{h'i} v_{h'j} \Ebb(w_i^2|X) \Ebb(w_j^2|X) \Bigr) & \leq \sum_{h,h' = 1}^{k} \sum_{i \neq j}^{n} \Ebb\Bigl( v_{hi} v_{hj} v_{h'i} v_{h'j} \Bigr) \\
	& \leq \sum_{h,h' = 1}^{k} \Ebb\Bigl( \sum_{i = 1}^{n}v_{h'i} v_{hi} \sum_{j = 1}^{n} v_{h'j} v_{hj} \Bigr) \\
	& = \sum_{h = 1}^{k} 1 = k.
	\end{align*}

	To bound the second term, we note that $\Var(w_i^4|X) \leq 1$, and along with \eqref{eqn: asmp_1} this yields
	\begin{align*}
	\sum_{h,h' = 1}^{k} \sum_{i = 1}^{n} \Ebb \left(v_{hi}^2 v_{h'i}^2 \Var(w_i^4|X)\right) \leq \sum_{h,h' = 1}^{n} \sum_{i = 1}^{n} \frac{c_1^4}{n^2} = \frac{c_1^4k^2}{n}.
	\end{align*}

	Finally, to bound the third term we note that if $h \neq h'$ and $i \neq i'$,
	$\mathrm{Cov} \bigl( v_{hi} ^2 \Ebb(w_i^2|X), v_{h'i'}^2 \Ebb(w_{i'}^2|X)\bigr) = 0$. Therefore,
	\begin{align*}
	\sum_{h,h' = 1}^{k} \sum_{i,i' = 1}^{n} \mathrm{Cov} \bigl( v_{hi} ^2 \Ebb(w_i^2|X), v_{h'i'}^2 \Ebb(w_{i'}^2|X)\bigr) & = \sum_{h = 1}^{k} \sum_{i = 1}^{n} \sum_{j = 1}^{n} \mathrm{Cov} \bigl( v_{hi} ^2 \Ebb(w_i^2|X), v_{hi'}^2 \Ebb(w_{i'}^2|X)\bigr) + \\
	& \quad \quad \sum_{h \neq h'}^{k} \sum_{i = 1}^{n} \mathrm{Cov} \bigl( v_{hi} ^2 \Ebb(w_i^2|X), v_{h'i}^2 \Ebb(w_{i}^2|X)\bigr) \\
	& \leq \sum_{h = 1}^{k} \sum_{i = 1}^{n} \sum_{j = 1}^{n} \frac{c_1^4}{n^2} + \sum_{h \neq h'}^{k} \sum_{i = 1}^{n} \frac{c_1^4}{n^2} \\
	& = c_1^4 k + c_1^4 \frac{k^2}{n},
	\end{align*}
	and putting the pieces together yields \eqref{eqn: variance_noise}.
\end{proof}

Now, we turn our attention to the third term of \eqref{eqn: LE_expansion}.
\begin{lemma}
	\begin{equation}
	\Ebb(w^T V_k V_k^T \thetast) = 0
	\end{equation}
	and assuming \eqref{eqn: asmp_1} holds,
	\begin{equation*}
	\Var(w^T V_k V_k^T \thetast) \leq k \frac{c_1^2 \delta^2(f,g)}{p_{\min}^2}
	\end{equation*}
\end{lemma}

Finally, we examine the first term in \eqref{eqn: LE_expansion}. We will need the following result.
\begin{lemma}
	For any $0 < \delta < 1$, there exists a number $c_2 > 0$ depending only on $d$ and $\delta$ such that, if $r \geq c_2(\log n/n)^{1/d}$, then
	\begin{equation*}
	\norm{D \thetast}^2 \leq \frac{16 L^4}{p_{\min}^2} n^2 r^{d + 2}
	\end{equation*}
	and
	\begin{equation*}
	\lambda_k \geq \frac{4 k^{2/d}}{\pi^2 n^{2/d}}
	\end{equation*}
	with probability at least $1 - \delta$.
\end{lemma}
\begin{lemma}
	\label{lem: approximation_error}
	For any $\delta > 0$, there exists a number $c_2 > 0$ constant in sample size such that if $r \geq c_2 (\log n / n)^{1/d}$,
	\begin{equation*}
	(\thetast)^T V_k V_k^T \thetast \geq n (\delta(f,g))^2 - \frac{n^2r^{d + 2} n^{2/d}}{k^{2/d}}
	\end{equation*}
	with probability at least $1 - \delta$.
\end{lemma}
\begin{proof}
	Letting $z = D \thetast$, we have
	\begin{align*}
	(\thetast)^T V_k V_k^T \thetast & = \norm{\thetast}^2 - (\thetast)^T \left(I - V_k V_k^T \right) \thetast \\
	& = \norm{\thetast}^2- z^T (D^{\dagger})^T (I - V_k V_k^T) D^{\dagger} z \\
	& \geq \norm{\thetast}^2 - z^T z \lambda_{\max} \Bigl((D^{\dagger})^T (I - V_k V_k^T) D^{\dagger} \Bigr) \\
	& \geq \norm{\thetast}^2 - \frac{z^T z}{\lambda_{k}}
	\end{align*}
	Examining $\norm{\thetast}^2$, we have that
	\begin{equation*}
	\Ebb(\norm{\thetast}^2) = n \delta^2(f,g) \quad \text{and} \quad \Var(\norm{\thetast}^2) \leq n \delta^2(f,g)
	\end{equation*}
	Turning to the second term, we have given our choice of $r$, along with the fact $f - g \in \mathcal{H}_1^{d}(2L)$,
	\begin{equation*}
	z^T z = \norm{D \thetast}^2 \leq \frac{16 L^4}{p_{\min}^2} n^2 r^{d + 2}
	\end{equation*}
	and additionally
	\begin{equation*}
	\lambda_k \geq \frac{4 k^{2/d}}{\pi^2 n^{2/d}},
	\end{equation*}
	each with probability $1 - \delta$.
\end{proof}

\end{document}

### 2019-06-19 notes/weekly_notes/6.18.19.NotesForWeek.tex
added weekly notes

--- 2019-06-11 notes/weekly_notes/6.9.19.NotesForWeek.tex
@@
 \newcommand{\dr}{\,dr}
+\newcommand{\dxpr}{\,dx'}
+\newcommand{\dypr}{\,dy'}
+\newcommand{\wt}[1]{\widetilde{#1}}
 
@@
 \begin{document}
-\title{Notes for Week 6/9/19 - 6/11/19}
+\title{Notes for Week 6/13/19 - 6/18/19}
 \author{Alden Green}
@@
 \end{equation*}
-with $f$,$g$ bounded away from $0$, $f(x), g(x) > p_{\min}$ for all $x \in [0,1]^d$.
+with $f$,$g$ bounded above and below, $0 < p_{\min} < f(x), g(x) < p_{\max} < \infty$ for all $x \in [0,1]^d$.
 
@@
 \end{equation*}
-We will consider those density functions which belong to $\mathcal{H} := \mathcal{H}_{1}^d(L)$, the $d$-dimensional Lipschitz ball with norm $L$.
+We will consider the function class $\mathcal{H} := \mathcal{H}_{1}^d(L)$, the $d$-dimensional Lipschitz ball with norm $L$.
 Formally, a function $f \in \mathcal{H}$ if for all $x,y \in [0,1]^d$, $\abs{f(x) - f(y)} \leq L \norm{x - y}$.\footnote{Note that the Lipschitz requirement, along with the restriction that $f$ be a density function over a bounded domain, imply that $f$ has finite $L^2$ norm.}
 
-\section{Sieve-based test.}
+\section{Test Statistic}
 
-For a radius $r > 0$ to be determined later, the $r$-neighborhood graph $G = (V,E)$ consists of vertices $V = [n]$ corresponding to the $n$ data points, and edges $E = \set{(u,v): \norm{x_u - x_v} \leq r}$.
-Write $D$ for the incidence matrix of this graph, and $L = D^T D$ for the graph Laplacian.
-Denote the singular value decomposition of $D$ by $D = U \Lambda^{1/2} V^T$, where $U$ and $V$ are orthonormal matrices and $\Lambda$ is a diagonal matrix with entries $0 = \lambda_1 \leq \lambda_2 \leq \lambda_3 \leq \ldots \leq \lambda_n$, so that $L = V \Lambda V^T$.
-For $k = 1,\ldots,n$, write $V_k = (v_1 \ldots v_k)$ for the $n \times k$ matrix containing the first $k$ columns of $V$--that is, the first $k$ eigenvectors of $L$.
-Our first test statistic will be as follows:
-
-\begin{equation}
-\label{eqn: laplacian_eigenmaps_statistic}
-T_{LE}^{(k)} := \sup_{\substack{\theta \in \mathrm{col}(V_k), \\ \norm{\theta} = 1} } \dotp{\theta}{\ell}^2 = \ell^T V_k V_k^T \ell.
-\end{equation}
-
-\subsection{Assumptions}
-The following are statements I wish to hold in order for the later theory to follow, but I cannot yet prove:
-\begin{enumerate}
-	\item There exists a constant $c_1 > 0$ -- which may depend on dimension $d$ but not on $n$ --  such that for all $h \in 1, \ldots,k$ and all $i = 1,\ldots,n$, 
-	\begin{equation}
-	\label{eqn: asmp_1}
-	\abs{v_{hi}} \leq \frac{c}{\sqrt{n}}
-	\end{equation}
-\end{enumerate}
-
-If $v_k$ were eigenvectors of the grid--rather than of the neighborhood graph--the first item would hold with $c = 2^{d/2}$.
-
-\subsection{Main result.}
-Let $d = 3,4$, choose $r = c(\log(n)/n)^{1/d}$ to satisfy the condition of Lemma \ref{lem: approximation_error}, and let $k = n^{2d/(d + 4)}$. Let 
+Let $r > 0$ be a bandwidth parameter, let $\eta_r: \mathcal{X} \times \mathcal{X} \to [0,\infty)$ be a Mercer kernel given by
 \begin{equation*}
-\phi_{LE} =
+\eta_r(x,y) =
 \begin{cases}
-1,~ & \text{if}~ T_{LE}^{(k)} \geq k + a\sqrt{2k + c^4\left(k + \frac{k^2}{n}\right)} \\
-0,~ & \text{otherwise}.
+1, \quad & \textrm{$\norm{x - y} \leq r$} \\
+0, \quad & \textrm{$\norm{x - y} > r$},
 \end{cases}
 \end{equation*}
-The following theorem holds.
-\begin{theorem}
-	There exists some number $c > 0$ constant in sample size $n$ such that for any $\epsilon \geq c n^{-2/(4 + d)}$, we have
+and let $\widetilde{A}$ be an $n \times n$ matrix with entries $\widetilde{A}_{ij} = n^{-1}\eta(x_i,x_j)$.
+We obtain $A$ from $\widetilde{A}$ be setting all diagonal entries of $\widetilde{A}$ to be zero.
+Our test statistic is then
 \begin{equation*}
-	R_{\epsilon}^{(t)}(\phi_{LE}; \Hclass_1^{d}(L)) \leq \frac{1}{a^2}.
-	\end{equation*}
-	when $d = 3$ or $4$. 
-\end{theorem}
-\begin{proof}
-
-\end{proof}
-
-\subsection{Type I error}
-
-Under the null hypothesis, $\ell$ consists of $n$ independent and identically distributed Rademacher random variables.
-Using this fact, we can bound the first two moments of $T_{LE}$ under the null hypothesis.
-
-\begin{lemma}
-	If $f = g$, then
-	\begin{equation}
-	\Ebb(T_{LE}^{(k)}) = k
-	\end{equation}
-	and
-	\begin{equation}
-	\label{eqn: variance_noise_typeI}
-	\Var(T_{LE}^{(k)}) \leq 2k + c_1^4\left(k + \frac{2k^2}{n}\right)
-	\end{equation}
-\end{lemma}
-\begin{proof}
-	Taking the expectation first, we have
-	\begin{align*}
-	\Ebb(T_{LE}^{(k)}) & = \sum_{j = 1}^{k} \Ebb((\ell^T v_j)^2) \\
-	& = \sum_{j = 1}^{k} \sum_{i = 1}^{n} \sum_{i' = 1}^{n} \Ebb(\ell_i \ell_{i'}) v_{ji} v_{ji'} \\
-	& =  \sum_{j = 1}^{k} \sum_{i = 1}^{n} v_{ji}^2 = k.
-	\end{align*}
-
-	We defer the proof of \eqref{eqn: variance_noise_typeI} until the following subsection, where we show it in greater generality.
-\end{proof}
-
-\subsection{Type II error.}
-
-Write $\ell =: \thetast + w$, where $\thetast = (\thetast_i)$ is defined by
-\begin{equation*}
-\thetast_i = \frac{f(x_i) - g(x_i)}{f(x_i) + g(x_i)}
-\end{equation*}
-and $w = (w_i)$ therefore consists of $n$ independent (although not identically distributed) zero-mean noise terms. We expand
-\begin{equation}
-\label{eqn: LE_expansion}
-\ell^T V_k V_k^T \ell = (\thetast)^T V_k V_k^T \thetast + w^T V_k V_k w + 2 w^T V_k V_k^T \thetast,
-\end{equation}
-and focus at first our attention on the second term in \eqref{eqn: LE_expansion}.
-
-\begin{lemma}
-	Assuming \eqref{eqn: asmp_1} holds,
-	\begin{equation}
-	\label{eqn: expectation_noise_typeII}
-	\Ebb \left( w^T V_k V_k w\right) \geq k - \frac{kc_1^2 \delta^2(f,g)}{p_{\min}^2}
-	\end{equation}
-	and
-	\begin{equation}
-	\label{eqn: variance_noise}
-	\Var(T_{LE}^{(k)}) \leq 2k + c_1^4\left(k + \frac{2k^2}{n}\right)
-	\end{equation}
-\end{lemma}
-\begin{proof}
-	We note that, conditional on $X$,
-	\begin{equation*}
-	w_i = 
-	\begin{cases}
-	2\frac{g(x_i)}{f(x_i) + g(x_i)},~ & \text{with probability}~ \frac{f(x_i)}{f(x_i) + g(x_i)} \\
-	-2\frac{f(x_i)}{f(x_i) + g(x_i)},~ & \text{with probability}~ \frac{g(x_i)}{f(x_i) + g(x_i)}
-	\end{cases}
+T_{\eta} := \ell^T A \ell.
 \end{equation*}
 
-	We first show \eqref{eqn: expectation_noise_typeII}. Expanding the quadratic form as a double sum, and using the law of iterated expectation, we have
-	\begin{align}
-	\label{eqn: expectation_noise_typeII_1}
-	\Ebb \left( w^T V_k V_k w\right) & = \sum_{h = 1}^{k} \sum_{i = 1}^{n} \sum_{j = 1}^n \Ebb \bigl(w_i w_j v_{hi} v_{hj} \bigr) \nonumber \\
-	& = \sum_{h = 1}^{k} \sum_{i = 1}^{n} \sum_{j = 1}^n \Ebb \biggl( \Ebb\Bigl(w_i w_j \bigl\vert x_i, x_j\Bigr) v_{hi} v_{hj}\biggr) \nonumber \\
-	& = \sum_{h = 1}^{k} \sum_{i = 1}^{n} \Ebb \biggl( \Ebb\Bigl(w_i^2 \bigl\vert x_i\Bigr) v_{hi}^2\biggr).
-	\end{align}
-	The conditional expectation $\Ebb\Bigl(w_i^2 \bigl\vert x_i\Bigr)$ can be directly computed,
-	\begin{align*}
-	\Ebb\Bigl(w_i^2 \bigl\vert x_i\Bigr) & = 4\left(\frac{g(x_i)}{f(x_i) + g(x_i)}\right)^2 \frac{f(x_i)}{f(x_i) + g(x_i)} + 4\left(\frac{f(x_i)}{f(x_i) + g(x_i)}\right)^2 \frac{g(x_i)}{f(x_i) + g(x_i)} \\
-	& = 4 \frac{f(x_i)g(x_i)(f(x_i) + g(x_i))}{(f(x_i) + g(x_i))^3} \\
-	& = 4 \frac{f(x_i)g(x_i)}{(f(x_i) + g(x_i))^2}.
-	\end{align*}
-	and plugging this back into \eqref{eqn: expectation_noise_typeII_1}, we obtain
-	\begin{align*}
-	\Ebb \left( w^T V_k V_k w\right) & = \sum_{h = 1}^{k} \sum_{i = 1}^{n} \Ebb \left(4 \frac{f(x_i) g(x_i)}{(f(x_i) + g(x_i))^2} v_{hi}^2\right) \\
-	& = \sum_{h = 1}^{k} \sum_{i = 1}^{n} \Ebb \left(v_{hi}^2 \left(1 - \left(\frac{f(x_i) - g(x_i)}{f(x_i) + g(x_i)} \right)^2 \right)\right) \\
-	& \geq k - \frac{c^2}{n}\sum_{h = 1}^{k} \sum_{i = 1}^{n} \Ebb \left(\frac{f(x_i) - g(x_i)}{f(x_i) + g(x_i)} \right)^2 \\
-	& \geq k - \frac{kc_1^2 \delta^2(f,g)}{p_{\min}^2}
-	\end{align*}
-	and \eqref{eqn: expectation_noise_typeII} is shown.
+\section{Theory}
 
-	We now show \eqref{eqn: variance_noise}. It will be helpful to note that for all $i,i',j,j' \in [n]$,
+\begin{theorem}
+
+\end{theorem}
+
+\section{Supporting Theory}
+
+We begin by recalling Chebyshev's inequality, which we will use to bound deviations of $T_{\eta}$ from its expectation.
+\begin{lemma}[Chebyshev]
+	\label{lem: chebyshev}
+	Let $X$ be a random variable with $\Ebb X < \infty$ and $0 < \Var(X) < \infty$. Then,
 	\begin{equation*}
-	\mathrm{Cov}(w_iw_j, w_{i'}w_{j'} | X) = 
-	\begin{cases}
-	0, & \text{if}~ \set{i,i',j,j'}~ \textrm{has four distinct elements} \\
-	0, & \text{if}~ \set{i,i',j,j'}~ \textrm{has three distinct elements} \\
-	0, & \text{if}~ \set{i,i',j,j'}~ \textrm{has two distinct elements and}~ i = j \\
-	\Ebb(w_i^2|X) \Ebb(w_j^2|X), & \text{if}~ \set{i,i',j,j'}~ \textrm{has two distinct elements and}~ i \neq j \\
-	\Var(w_i^4|X), & \text{if}~ \set{i,i',j,j'}~ \textrm{has one distinct element.}
-	\end{cases}
-	\end{equation*}
-	and additionally
-	\begin{equation*}
-	\Ebb(w_iw_j|X) =
-	\begin{cases}
-	0, & \text{if}~ i \neq j \\
-	E(w_i^2|X), & \text{if}~ i = j.
-	\end{cases}
-	\end{equation*}
-	Then, by the law of total covariance,
-	\begin{align*}
-	\mathrm{Var}(w^T V_k V_k^T w) & = \sum_{h,h' = 1}^{k} \sum_{i,i' = 1}^{n} \sum_{j,j' = 1}^{n} \mathrm{Cov} \bigl(w_iw_jv_{hi} v_{hj}, w_{i'}w_{j'}v_{h'i'} v_{h'j'}\bigr) \\
-	& = \sum_{h,h' = 1}^{k} \sum_{i,i' = 1}^{n} \sum_{j,j' = 1}^{n} \Ebb \Bigl(  v_{hi} v_{hj} v_{h'i'} v_{h'j'} \mathrm{Cov} \bigl(w_iw_j, w_{i'}w_{j'} |X\bigr)\Bigr) + \\
-	& \quad \quad \mathrm{Cov} \Bigl( v_{hi}v_{hj} \Ebb(w_iw_j|X), v_{h'i'}v_{h'j'} \Ebb(w_i'w_j'|X)  \Bigr) \\
-	& = 2 \sum_{h,h' = 1}^{k} \sum_{i \neq j}^{n} \Ebb\Bigl( v_{hi} v_{hj} v_{h'i} v_{h'j} \Ebb(w_i^2|X) \Ebb(w_j^2|X) \Bigr) + \\
-	& \quad \quad \sum_{h,h' = 1}^{k} \sum_{i = 1}^{n} \Ebb \left(v_{hi}^2 v_{h'i}^2 \Var(w_i^4|X)\right) +  \\
-	& \quad \quad \quad \sum_{h,h' = 1}^{k} \sum_{i,i' = 1}^{n} \mathrm{Cov} \bigl( v_{hi} ^2 \Ebb(w_i^2|X), v_{h'i'}^2 \Ebb(w_{i'}^2|X)\bigr)
-	\end{align*}
-	Then, using the upper bound $\Ebb(w_i^2|X),\Ebb(w_j^2|X) \leq 1$, we can bound the first term
-	\begin{align*}
-	\sum_{h,h' = 1}^{k} \sum_{i \neq j}^{n} \Ebb\Bigl( v_{hi} v_{hj} v_{h'i} v_{h'j} \Ebb(w_i^2|X) \Ebb(w_j^2|X) \Bigr) & \leq \sum_{h,h' = 1}^{k} \sum_{i \neq j}^{n} \Ebb\Bigl( v_{hi} v_{hj} v_{h'i} v_{h'j} \Bigr) \\
-	& \leq \sum_{h,h' = 1}^{k} \Ebb\Bigl( \sum_{i = 1}^{n}v_{h'i} v_{hi} \sum_{j = 1}^{n} v_{h'j} v_{hj} \Bigr) \\
-	& = \sum_{h = 1}^{k} 1 = k.
-	\end{align*}
-
-	To bound the second term, we note that $\Var(w_i^4|X) \leq 1$, and along with \eqref{eqn: asmp_1} this yields
-	\begin{align*}
-	\sum_{h,h' = 1}^{k} \sum_{i = 1}^{n} \Ebb \left(v_{hi}^2 v_{h'i}^2 \Var(w_i^4|X)\right) \leq \sum_{h,h' = 1}^{n} \sum_{i = 1}^{n} \frac{c_1^4}{n^2} = \frac{c_1^4k^2}{n}.
-	\end{align*}
-
-	Finally, to bound the third term we note that if $h \neq h'$ and $i \neq i'$,
-	$\mathrm{Cov} \bigl( v_{hi} ^2 \Ebb(w_i^2|X), v_{h'i'}^2 \Ebb(w_{i'}^2|X)\bigr) = 0$. Therefore,
-	\begin{align*}
-	\sum_{h,h' = 1}^{k} \sum_{i,i' = 1}^{n} \mathrm{Cov} \bigl( v_{hi} ^2 \Ebb(w_i^2|X), v_{h'i'}^2 \Ebb(w_{i'}^2|X)\bigr) & = \sum_{h = 1}^{k} \sum_{i = 1}^{n} \sum_{j = 1}^{n} \mathrm{Cov} \bigl( v_{hi} ^2 \Ebb(w_i^2|X), v_{hi'}^2 \Ebb(w_{i'}^2|X)\bigr) + \\
-	& \quad \quad \sum_{h \neq h'}^{k} \sum_{i = 1}^{n} \mathrm{Cov} \bigl( v_{hi} ^2 \Ebb(w_i^2|X), v_{h'i}^2 \Ebb(w_{i}^2|X)\bigr) \\
-	& \leq \sum_{h = 1}^{k} \sum_{i = 1}^{n} \sum_{j = 1}^{n} \frac{c_1^4}{n^2} + \sum_{h \neq h'}^{k} \sum_{i = 1}^{n} \frac{c_1^4}{n^2} \\
-	& = c_1^4 k + c_1^4 \frac{k^2}{n},
-	\end{align*}
-	and putting the pieces together yields \eqref{eqn: variance_noise}.
-\end{proof}
-
-Now, we turn our attention to the third term of \eqref{eqn: LE_expansion}.
-\begin{lemma}
-	\begin{equation}
-	\Ebb(w^T V_k V_k^T \thetast) = 0
-	\end{equation}
-	and assuming \eqref{eqn: asmp_1} holds,
-	\begin{equation*}
-	\Var(w^T V_k V_k^T \thetast) \leq k \frac{c_1^2 \delta^2(f,g)}{p_{\min}^2}
+	\Pbb(\abs{X - \Ebb X} \geq a) \leq \frac{\Var(X)}{a^2}.
 	\end{equation*}
@@
 
-Finally, we examine the first term in \eqref{eqn: LE_expansion}. We will need the following result.
+In order to employ Lemma \ref{lem: chebyshev}, we must bound the expectation and variance of $T_{\eta}$ under the null and alternative hypotheses. We begin with the expectation $\Ebb T_{\eta}$. 
 \begin{lemma}
-	For any $0 < \delta < 1$, there exists a number $c_2 > 0$ depending only on $d$ and $\delta$ such that, if $r \geq c_2(\log n/n)^{1/d}$, then
+	\label{lem: mmd_expectation}
+	Under the null hypothesis $h = g$, for any $h, g \in \mathcal{H}_{1}^{d}(L)$,
 	\begin{equation*}
-	\norm{D \thetast}^2 \leq \frac{16 L^4}{p_{\min}^2} n^2 r^{d + 2}
+	\Ebb T_{\eta} = 0.
 	\end{equation*}
-	and
+
+	Under the alternative hypothesis $h \neq g$, for any $h, g \in \mathcal{H}_{1}^{d}(L)$, 
 	\begin{equation*}
-	\lambda_k \geq \frac{4 k^{2/d}}{\pi^2 n^{2/d}}
+	\Ebb T_{\eta} \geq \frac{(n - 1)p_{\min}^2}{ p_{\max}^2} \left(\delta^2(f,g)\frac{r^d}{2^d} - L\delta(f,g)r^{d + 1}\right)
 	\end{equation*}
-	with probability at least $1 - \delta$.
-\end{lemma}
-\begin{lemma}
-	\label{lem: approximation_error}
-	For any $\delta > 0$, there exists a number $c_2 > 0$ constant in sample size such that if $r \geq c_2 (\log n / n)^{1/d}$,
-	\begin{equation*}
-	(\thetast)^T V_k V_k^T \thetast \geq n (\delta(f,g))^2 - \frac{n^2r^{d + 2} n^{2/d}}{k^{2/d}}
-	\end{equation*}
-	with probability at least $1 - \delta$.
 \end{lemma}
 \begin{proof}
-	Letting $z = D \thetast$, we have
+	Let $\thetast$ be a vector in $\Reals^{n}$, with entries $ \thetast_i = \Ebb(\ell_i|x_i)$.
+Letting $w = \ell - \thetast$, we note that conditional on $X$, the noise component $w$ consists of $n$ independent and zero mean terms.
+
+	We expand the test statistic
+	\begin{equation*}
+	T_{\eta} = (\thetast)^T A \thetast + w^T A w + 2 w^T A \thetast
+	\end{equation*}
+	and note that by the law of iterated expectation, 
+	\begin{equation*}
+	\Ebb(T_{\eta}) = \Ebb(T_{\eta}|X) = \Ebb\left( (\thetast)^T A \thetast \right)
+	\end{equation*}
+
+	Now, let $X$ and $Y$ be independent random variables with distribution $\mu$, and let $\Delta(x) := f(x) - g(x)$. By the linearity of expectation,
 	\begin{align*}
-	(\thetast)^T V_k V_k^T \thetast & = \norm{\thetast}^2 - (\thetast)^T \left(I - V_k V_k^T \right) \thetast \\
-	& = \norm{\thetast}^2- z^T (D^{\dagger})^T (I - V_k V_k^T) D^{\dagger} z \\
-	& \geq \norm{\thetast}^2 - z^T z \lambda_{\max} \Bigl((D^{\dagger})^T (I - V_k V_k^T) D^{\dagger} \Bigr) \\
-	& \geq \norm{\thetast}^2 - \frac{z^T z}{\lambda_{k}}
+	\Ebb (\thetast)^T A \thetast & = \sum_{i \neq j = 1}^{n} \Ebb( \thetast_i \thetast_j A_{ij}) \\
+	& = (n - 1) \Ebb \bigl(\thetast_1 \thetast_2 \eta_r(x_1, x_2) \bigr) \\
+	& \geq \frac{4(n - 1)}{p_{\max}^2} \Ebb(\Delta(x) \Delta(y) \eta_r(X,Y))
 	\end{align*}
-	Examining $\norm{\thetast}^2$, we have that
+
+	We can lower bound $\Ebb(\Delta(x) \Delta(y) \eta_r(X,Y))$ by $\delta^2(f,g)$, minus a remainder term, as follows
+	\begin{align*}
+	\Ebb(\Delta(x) \Delta(y) \eta_r(X,Y)) & = \iint \Delta(x) \Delta(y) \eta_r(x,y) d\mu(x) d\mu(x) \\
+	& \geq \frac{p_{\min}^2}{4} \iint \Delta(x) \Delta(y) \eta_r(x,y) dx dy \\
+	& \geq \frac{p_{\min}^2}{4} \left( \iint \Delta^2(x) \eta_r(x,y) dx dy - \iint \abs{\Delta(x)} \abs{\Delta(x) - \Delta(y)} \eta(x,y) dx dy \right)
+	\end{align*}
+	To lower bound the first term, observe that for any $x \in [0,1]^d$,
 	\begin{equation*}
-	\Ebb(\norm{\thetast}^2) = n \delta^2(f,g) \quad \text{and} \quad \Var(\norm{\thetast}^2) \leq n \delta^2(f,g)
+	\int_{[0,1]^d} \eta_r(x,y) dx \geq \frac{r^d}{2^d}
 	\end{equation*}
-	Turning to the second term, we have given our choice of $r$, along with the fact $f - g \in \mathcal{H}_1^{d}(2L)$,
+	and therefore
 	\begin{equation*}
-	z^T z = \norm{D \thetast}^2 \leq \frac{16 L^4}{p_{\min}^2} n^2 r^{d + 2}
+	\iint \Delta^2(x) \eta_r(x,y) dx dy \geq \frac{r^d}{2^d} \int_{[0,1]^d} \Delta^2(x) dx \geq \frac{r^d}{2^d} \delta^2(f,g).
 	\end{equation*}
-	and additionally
+
+	To upper bound the remainder term, observe that $\Delta \in \mathcal{H}_{1}^{d}(2L)$. Using this fact, we obtain the upper bound
+	\begin{align*}
+	\iint \abs{\Delta(x)} \abs{\Delta(x) - \Delta(y)} \eta(x,y) dx dy & \leq : \iint \abs{\Delta(x)} \norm{x - y} \eta(x,y) dx dy \\
+	& \leq L r^{d + 1} \int_{[0,1]^d} \abs{\Delta(x)} dx \\
+	& \leq L r^{d + 1} \delta(f,g).
+	\end{align*}
+	where the last inequality follows from Jensen's inequality.
+\end{proof}
+
+\begin{lemma}
+	\label{lem: var_mmd}
+	For any $h,g \in \mathcal{H}_1^{d}(L)$,
 	\begin{equation*}
-	\lambda_k \geq \frac{4 k^{2/d}}{\pi^2 n^{2/d}},
+	\Var(T_{\eta}) \leq 2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + 2r^d
 	\end{equation*}
-	each with probability $1 - \delta$.
+\end{lemma}
+\begin{proof}
+	We begin by computing the second moment $\Ebb T_{\eta}^2$. Expanding the quadratic form, we obtain
+	\begin{align}
+	\Ebb T_{\eta}^2 =  \sum_{i,i' = 1}^{n} \sum_{j,j' = 1}^{n} \Ebb(\ell_i \ell_{i'} \ell_{j} \ell_{j'} A_{ii'} A_{jj'})
+	\end{align}
+
+	We next compute this expectation, based on the number of distinct indices $i,i',j,j'$.
+Let $X, X', Y, Y'$ be four independent random variables, each with distribution $\mu$, and write $\widetilde{\Delta} := 2 \Delta / \mu$.
+Then, applying the law of conditional expectation, we obtain
+	\begin{equation*}
+	n^2 \Ebb(\ell_i \ell_{i'} \ell_{j} \ell_{j'} A_{ii'} A_{jj'}) = 
+	\begin{cases}
+	\Ebb(\widetilde{\Delta}(X) \wt{\Delta}(X') \eta_r(X,X'))^2, \quad & \textrm{if}~ \set{i,i',j,j'} ~\textrm{contains four distinct elements}, \\
+	\Ebb(\widetilde{\Delta}(X') \wt{\Delta}(Y') \eta_r(X,X') \eta_r(X,Y')), \quad & \textrm{if}~ \set{i,i',j,j'} ~\textrm{contains three distinct elements and}~ i = j, \\
+	\Ebb\biggl(\eta_r(X,Y)^2\biggr), \quad  & \textrm{if}~ \set{i,i',j,j'} ~\textrm{contains two distinct elements and}~ i = j. \\
+	0, \quad & \textrm{if}~ i = i' ~\textrm{or}~ j = j'.
+	\end{cases}
+	\end{equation*}
+	and as a result we obtain
+	\begin{align}
+	\label{eqn: var_mmd_1}
+	n^2 \Ebb T_{\eta}^2 = & n(n-1)(n-2)(n-3)\Ebb\biggl(\widetilde{\Delta}(X) \wt{\Delta}(X') \eta_r(X,X')\biggr) + \nonumber \\
+	& \quad 2n(n-1)(n-2) \Ebb \biggl(\widetilde{\Delta}(X') \wt{\Delta}(Y') \eta_r(X,X') \eta_r(X,Y') \biggr) + \nonumber \\
+	& \quad \quad 2n(n-1) \Ebb\biggl(\eta_r(X,Y)^2\biggr).
+	\end{align}
+
+	The first term -- covering the case when $\set{i,i',j,j'}$ contains four distinct elements --  will be canceled out by the mean, as we will see, and we do not simplify it further.
+
+	The second term we upper bound by methods similar to those used to obtain the upper bound in Lemma \ref{lem: mmd_expectation}. We will need the following upper bound,
+	\begin{align*}
+	\int \eta(x,x') \eta(x,y') \dx & \leq r^d \1(\norm{x' - y'} \leq 2r) \\
+	& = r^d \eta_{2r}(x',y').
+	\end{align*}
+	We proceed,
+	\begin{align}
+	\label{eqn: var_mmd_2}
+	\Ebb \biggl(\widetilde{\Delta}(X') \wt{\Delta}(Y') \eta_r(X,X') \eta_r(X,Y') \biggr) & \leq \frac{p_{\max}^2}{p_{\min}^2} \iiint \Delta(x') \Delta(y') \eta(x,x') \eta(x,y') \dx \dypr \dxpr \nonumber\\
+	& \leq \frac{p_{\max}^2}{p_{\min}^2} r^d \iint \Delta(x') \Delta(y') \eta_{2r}(x',y') \dypr \dxpr \nonumber \\
+	& \leq \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right).
+	\end{align}
+
+	The following upper bound on the third term is a straightforward consequence of the fact $\eta_r^2 = \eta_r$, and therefore
+	\begin{equation}
+	\label{eqn: var_mmd_3}
+	\Ebb(\eta_r(X,Y)^2) = \Ebb(\eta_r(X,Y)) = r^d. 
+	\end{equation}
+
+	Finally, we have that
+	\begin{equation}
+	\label{eqn: var_mmd_4}
+	n^2 \Ebb(T_{\eta}^2) = \left(n(n-1) \Ebb(\widetilde{\Delta}(X) \wt{\Delta}(X') \eta_r) \right)^2 .
+	\end{equation}
+
+	Combining \eqref{eqn: var_mmd_1}-\eqref{eqn: var_mmd_4}, we obtain
+	\begin{equation*}
+	\Var(T_\eta) \leq 2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + 2r^d.
+	\end{equation*}
 \end{proof}

### 2019-07-16 notes/weekly_notes/7.10.19.NotesForWeek.tex
weekly notes

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbb{I}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 7/10/19 - 7/16/19}
\author{Alden Green}
\date{\today}
\maketitle

\section{Testing over graphs.}

Suppose we observe $G = (V,E)$, an undirected graph over $V = [n]$. Let $D$ be the $m \times n$ incidence matrix of $G$, with singular value decomposition $D = U\Lambda^{1/2}V^T$, so that the Laplacian matrix $L = V \Lambda V^T$. For $i \in [n]$, we observe
\begin{equation*}
z_i = \beta_i + \varepsilon_i, \quad \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
\end{equation*}
Our statistical goal is hypothesis testing. We wish to distinguish
\begin{equation*}
\mathbf{H}_0: \norm{\beta}_2 = 0 \quad \textrm{vs.} \quad \mathbf{H}_a: \norm{\beta}_2 > 0.
\end{equation*}
We will evaluate our performance using the notion of \emph{worst-case error}. For a given ``function'' class $\mathcal{H}$, test function $\phi: \Reals^n \to \set{0,1}$, and $\epsilon > 0$, let
\begin{equation*}
\mathcal{R}_{\epsilon}(\phi;\mathcal{H})= \Ebb_{0}(\phi) + \sup_{\beta \in \mathcal{H}:\norm{\beta}_2 > \epsilon} \Ebb_{\beta}(1 - \phi)
\end{equation*}

An example function class we will consider will be unit balls in discrete Sobolev norms, with known smoothness. For $s,d$ positive integers, and radius $C_n > 0$, let
\begin{equation*}
\mathcal{S}^{s}_{d}(C_n) = \set{\beta: \norm{D^{(s)}\beta}_2 \leq C_n}
\end{equation*}
where
\begin{equation*}
D^{(s)} = 
\begin{cases}
L^{s/2}, & \text{$s$ even} \\
DL^{(s-1)/2}, & \text{$s$ odd}.
\end{cases}
\end{equation*}
We note that the constraint is equivalent to $\beta^T V \Lambda^s V \beta \leq C_n^2$.
\subsection{Test statistic}
Let $0 = \lambda_0 \leq \lambda_1 \leq \lambda_2 \leq \ldots \leq \lambda_{n - 1}$ denote the ordered eigenvalues of $L$, and let $v_k$ denote the eigenvector corresponding to $\lambda_k$. For a constant $C > 0$ to be specified later, let 
\begin{equation*}
T_C = \sum_{k: \lambda_k^{s} \leq C^2} y_k^2
\end{equation*}
where
\begin{equation*}
y_k = \frac{1}{\sqrt{n}} \dotp{z}{v_k}.
\end{equation*}

We first compute the expectation $\Ebb(T_C)$. Let $\theta \in \Reals^n$ represent the expectation of $(y_k)$, meaning
\begin{equation*}
\theta_k = \frac{1}{\sqrt{n}}\dotp{\beta}{v_k}
\end{equation*}
and let $\Pi_C\theta$ have entries $(\Pi_C \theta)_k = \1(\lambda_k^{s} \leq C^2) \theta_k$. Finally, let $N(C) = \sharp\set{k: \lambda_k^{s} \leq C^2}$.

\begin{lemma}
	\label{lem:expectation}
	For any $\beta \in \Reals^n$,
	\begin{equation}
	\label{eqn:expectation_1}
	\Ebb(T_C) = \frac{N(C)}{n} + \norm{\Pi_C \theta}_2^2 
	\end{equation}
	If additionally $\beta \in \mathcal{H}$, the following lower bound holds:
	\begin{equation}
	\label{eqn:expectation_2}
	\Ebb(T_C) \geq \frac{N(C)}{n} + \frac{\norm{\beta}^2}{n}  - \frac{C_n^2}{nC^{2}}
	\end{equation}
\end{lemma}
\begin{proof}
	We can write
	\begin{align*}
	\Ebb(T_C) & = \sum_{k: \lambda_k^s \leq C} \Ebb(y_k^2) \\
	& = \frac{1}{n}\sum_{k: \lambda_k^s \leq C} \Ebb\bigl(\dotp{\beta}{v_k}^2 + \dotp{\varepsilon}{v_k}^2 + 2 \dotp{\varepsilon}{v_k} \dotp{\beta}{v_k}\bigr) \\
	& = \sum_{k: \lambda_k^s \leq C}\theta_k^2 + \frac{1}{n} \\
	& = \norm{\Pi_C \theta}^2 + \frac{N(C)}{n},
	\end{align*}
	showing \eqref{eqn:expectation_1}. Now, assuming, $\norm{D^{(s)}\beta}_2 \leq C_n$, we can further obtain
	\begin{align*}
	\norm{\Pi_C \theta}^2 & = \norm{\theta}^2 - \sum_{k:\lambda_k^s > C^2} \theta_k^2 \\
	& \geq \norm{\theta}^2 - \frac{1}{C^{2}}\sum_{k:\lambda_k^s > C^2} \theta_k^2 \lambda_k^s \\
	& \geq \norm{\theta}^2 -\frac{1}{nC^2} \beta^T V \Lambda^s V^T \beta \\
	& \geq \norm{\theta}^2  - \frac{C_n^2}{nC^{2}}
	\end{align*}
	and \eqref{eqn:expectation_2} is shown.
\end{proof}

We now turn to computing the variance $\Var(T_C)$.
\begin{lemma}
	\label{lem:variance}
	\begin{equation*}
	\Var(T_C) = \frac{2N(C)}{n^2} + \frac{4 \norm{\Pi_C \theta}_2^2}{n}
	\end{equation*}
\end{lemma}
\begin{proof}
	To begin, we rewrite
	\begin{align*}
	T_C & = \sum_{k:\lambda_k^s \leq C^2} y_k^2 \\
	& = \frac{1}{n} \sum_{k: \lambda_k^s \leq C^2} \dotp{z}{v_k}^2 \\
	& = \frac{1}{n} \sum_{k: \lambda_k^s \leq C^2} z^T v_k v_k^T z \\
	& =: \frac{1}{n} z^T P_{C} z.
	\end{align*}
	where $P_C := \sum_{k: \lambda_k^s \leq C^2} v_k v_k^T$. Therefore $\Var(T_C) = \Var(z^T P_{C} z)/n^2$. We expand $z = \beta + \varepsilon$ to obtain
	\begin{align}
	\Var(z^T P_{C} z) & = \Var((\beta + \epsilon)^T P_{C} (\beta + \epsilon)) \nonumber \\
	& = \Var(\beta^T P_{C} \beta + 2 \epsilon P_{C} \beta + \epsilon^T P_{C} \epsilon) \nonumber \\
	& = 4 \beta^T P_{C} I P_{C} \beta + \Var(\epsilon^T P_{C} \epsilon) + 4 \Cov(\epsilon P_{C} \beta, \epsilon^T P_{C} \epsilon) \nonumber \\
	& = 4 n \norm{\Pi_C \theta}_2^2 + \Var(\epsilon^T P_{C} \epsilon) \label{eqn:variance_1}
	\end{align}
	where the last equality follows from the Gaussianity of $\epsilon$, as
	\begin{equation*}
	\Ebb\bigl( (\epsilon P_{C} \beta) (\epsilon^T P_{C} \epsilon) \bigr) = \sum_{i = 1}^{n} \sum_{j = 1}^{n} \sum_{k = 1}^{n} (P_C\beta)_k (P_C)_{ij} \Ebb(\epsilon_i \epsilon_j \epsilon_k) = 0.
	\end{equation*}
	Also by the Gaussianity of $\epsilon$, $\epsilon^T P_C \epsilon \sim \chi_{N(C)}^2$, and therefore $\Var(\epsilon^T P_{C} \epsilon) = 2N(C)$. Plugging back into \eqref{eqn:variance_1}, we obtain
	\begin{equation*}
	\Var(z^T P_{C} z) = 4 n \norm{\Pi_C \theta}_2^2 + 2N(C) 
	\end{equation*}
	and therefore the desired result is proved.
\end{proof}

We will consider now the test $\phi_C = \mathbf{1}\{T(C) \geq N(C)/n + \tau(b)\}$, where for $b \geq 1$, $\tau(b) = b \sqrt{2N(C)/n^2}$. We first upper bound the type I error.

\begin{lemma}
	\label{lem:type_I_error}
	Under the null hypothesis $\beta = 0$, and for any $C > 0$,
	\begin{equation*}
	\Ebb_{\beta = 0}(\phi_C) \leq \frac{1}{b^2}. 
	\end{equation*}
\end{lemma}
\begin{proof}
	The desired result follows from Chebyshev's inequality,
	\begin{align*}
	\Ebb_{\beta = 0}(\phi) & = \Pbb_{\beta = 0}\bigl(T(C) \geq N(C)/n + \tau(b)\bigr) \\ 
	& = \Pbb_{\beta = 0}\bigl(T(C) - \frac{N(C)}{n} \geq \tau(b)\bigr) \\
	& \leq \Pbb_{\beta = 0}\bigl(\abs{T(C) - \frac{N(C)}{n}} \geq \tau(b)\bigr) \\
	& \leq \frac{\Var_{\beta = 0}(T_C)}{\tau(b)^2} = \frac{1}{b^2}.
	\end{align*}
\end{proof}
The calculation for the type II error will be slightly more involved.
\begin{lemma}
	\label{lem:type_II}
	Let $b \geq 1$ be fixed. For every $\beta \in \mathcal{S}_d^s$ such that
	\begin{equation}
	\label{eqn:type_II}
	\frac{\norm{\beta}^2}{n} \geq 2b\sqrt{2\frac{N(C)}{n^2}} + \frac{C_n^2}{nC^2}
	\end{equation}
	we have that
	\begin{equation*}
	\Ebb_{\beta}(1 - \phi) \leq \frac{2}{b^2} + \frac{2}{b\sqrt{2N(C)}}.
	\end{equation*}
\end{lemma}
\begin{proof}
	Let $\Delta = \Ebb_{\beta}(T_C) - N(C)/n = \norm{\Pi_C \theta}^2$, and observe that by Lemma \ref{lem:expectation} and \eqref{eqn:type_II},
	\begin{equation*}
	\Delta \geq \frac{\norm{\beta}_2^2}{n} - \frac{C_n^2}{nC^2} \geq 2 \tau(b).
	\end{equation*}
	An application of Chebyshev's inequality yields
	\begin{align*}
	\Ebb_{\beta}\bigl(1 - \phi\bigr) & = \Pbb_{\beta}\bigl(T_C \leq N(C)/n + \tau(b)\bigr) \\
	& = \Pbb_{\beta}\bigl(T_C - \Ebb_{\beta}(T_C) \leq \tau(b) - \Delta \bigr) \\
	& \leq \Pbb_{\beta}\bigl(\abs{T_C - \Ebb_{\beta}(T_C)} \leq \Delta - \tau(b) \bigr) \tag{since $\Delta \geq \tau(b)$}	\\
	& \leq \frac{\Var_{\beta}(T_C)}{(\Delta - \tau(b))^2} \\
	& \leq 4\frac{\Var_{\beta}(T_C)}{\Delta^2} \tag{since $\Delta \geq 2\tau(b)$} \\
	& \leq 4\frac{2N(C)/n^2 + \norm{\Pi_C\theta}_2^2/n}{\Delta^2}.
	\end{align*}

	We now handle each summand separately. For the first term, since $\Delta \geq 2 \tau(b)$, we have
	\begin{equation*}
	\frac{2N(C)}{n^2\Delta^2} \leq \frac{1}{2b^2}.
	\end{equation*}

	For the second term, since $\Delta = \norm{\Pi_C\theta}^2$, we have
	\begin{align*}
	\frac{\norm{\Pi_C\theta}_2^2/n}{\Delta^2} & \leq \frac{1}{n\Delta^2} \\
	& \leq \frac{1}{2n\tau(b)} \\
	& = \frac{1}{2b\sqrt{2N(C)}}.
	\end{align*}
\end{proof}

To more explicitly specify the critical radius $\epsilon: \norm{\beta}_2 \geq \epsilon$, we will need to make an assumption on the relation between $N(C)$ and $C$. In particular, let $C = C^*$, where
\begin{equation*}
C^* = \frac{(C_n n^{s/d})^{4s/(4s + d)}}{n^{s/d}}.
\end{equation*}
We will assume the following bounds on $N(C^*)$:
\begin{enumerate}[label=(A\arabic*)]
	\item Tail decay:
	\label{asmp:tail_decay}
	\begin{equation*}
	N(C^*) \leq (C^*)^{d/s}n
	\end{equation*}
	\item Asymptotic consistency:
	\label{asmp:asymp_consistency}
	\begin{equation*}
	\lim_{n \to \infty} N(C^*) = \infty
	\end{equation*}
\end{enumerate}

\begin{corollary}
	\label{cor:critical_radius}
	Under \ref{asmp:asymp_consistency} and \ref{asmp:tail_decay}, 
	letting
	\begin{equation*}
	\epsilon^2 = (2\sqrt{2}b + 1) \left(\frac{(C_n n^{s/d})^{2d/(4s + d)}}{n}\right)
	\end{equation*}
	we have that
	\begin{equation*}
	\mathcal{R}_{\epsilon}(\phi_{C^*};\mathcal{S}_d^s) \leq \frac{2}{b^2} + o(1)
	\end{equation*}
\end{corollary}
\begin{proof}
	Recall that
	\begin{equation*}
	\mathcal{R}_{\epsilon}(\phi_{C^*};\mathcal{S}_d^s) = \Ebb_{\beta = 0}(\phi_{C^*}) + \sup_{\beta \in \mathcal{H}:\norm{\beta}_2/n > \epsilon} \Ebb_{\beta}(1 - \phi_{C^*})
	\end{equation*} 
	By Lemma \ref{lem:type_I_error}, we have that
	\begin{equation*}
	\Ebb_{\beta = 0}(\phi_{C^*}) \leq \frac{1}{b^2}.
	\end{equation*}
	Now, we verify that $\epsilon^2 \geq 2b\sqrt{2N(C^*)/n^2} + C_n^2/n(C^*)^2$. By Assumption \ref{asmp:tail_decay} and the choice of $C^{\star}$, we have
	\begin{equation*}
	N(C^*) \leq (C^*)^{d/s}n = (C_n n^{s/d})^{4d/(4s + d)}
	\end{equation*}
	and therefore
	\begin{equation*}
	2b\sqrt{\frac{N(C^*)}{n^2}} \leq \frac{2b}{n} (C_n n^{s/d})^{2d/(4s+d)}
	\end{equation*}
	Moving on to the second term, we have
	\begin{align*}
	\frac{C_n^2}{n(C^{\star})^2} & = \frac{C_n^2n^{2s/d}}{n(C_nn^{s/d})^{8s/(4s+d)}} \\
	& = \frac{C_n^{2d/(4s + d)}n^{(s/d)2d/(4s + d)}}{n} \\
	& = \frac{(C_n n^{s/d})^{2d/(4s + d)}}{n}
	\end{align*}
	and therefore $\epsilon \geq 2b\sqrt{2N(C^*)/n^2} + C_n^2/n(C^*)^2$. As a result, by Lemma \ref{lem:type_II} for any $\beta \in \mathcal{S}^s_d$ such that $\norm{\beta}/n \geq \epsilon$, 
	\begin{equation*}
	\Ebb_{\beta}(1 - \phi_{C^{\star}}) \geq \frac{2}{b^2} + \frac{2}{b\sqrt{2N(C)}}
	\end{equation*}
	and by Assumption \ref{asmp:asymp_consistency}, the latter term tends to infinity with $n$.
\end{proof}

\section{Applications}

\subsection{Grid}
Let $G$ be the $d$-dimensional grid graph, and let $C_n = n^{1/2 - s/d}$, so that
\begin{equation*}
C^{\star} = \frac{n^{2s/(4s + d)}}{n^{s/d}}
\end{equation*}
The following bound holds for eigenvalues of the Laplacian $L$ of the grid graph:
\begin{equation*}
\lambda_k^s \geq 4\sin^{2s}(\pi k^{1/d}/(2n^{1/d})) \geq \frac{\pi^{2s} k^{2s/d}}{4^{s}n^{2s/d}}
\end{equation*}
Therefore, for any $C > 0$, if $\lambda_k^s < C^2$, then
\begin{align*}
\frac{\pi^{2s} k^{2s/d}}{4^sn^{2s/d}} & < C^2 \Longrightarrow \\
k & < \frac{4^s C^{d/s}n}{\pi^{2s}}
\end{align*}
Since this holds in particular with respect to $C = C^{\star}$, assumption \ref{asmp:tail_decay} is satisfied. One can similarly shown \ref{asmp:asymp_consistency} holds as well. By Corollary \ref{cor:critical_radius}, we therefore have that when
\begin{equation*}
\epsilon^2 = (2b + 1)\left(\frac{n^{d/(4s + d)}}{n}\right) = (2b + 1)n^{-4s/(4s + d)}
\end{equation*}
we have
\begin{equation*}
\mathcal{R}_{\epsilon}(\phi_{C^*};\mathcal{S}_d^s) \leq \frac{2}{b^2} + o(1).
\end{equation*}

\section{Additional Theory}
We consider now a naive test statistic,
\begin{equation*}
T = \frac{\norm{z}^2}{n}
\end{equation*}
with associated test $\Phi_b(T) = \mathbf{1}(T \leq 1 + 2b/\sqrt{n})$.
and prove an upper bound showing that when $\epsilon \geq n^{-1/4}$, the worst-case error is bounded for all $\beta \in L^2(V)$.

\begin{lemma}
	Let $\epsilon = n^{-1/4}$. Then
	\begin{equation*}
	\mathcal{R}_{\epsilon}(\Phi_b;L^2(V)) \leq \frac{2}{b^2} + \frac{1}{n}
	\end{equation*}
\end{lemma}

\end{document}

### 2019-07-26 7.23.19.FutureDirections.tex
added future directions

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbb{I}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Future Directions}
\author{Alden Green}
\date{\today}
\maketitle

Throughout, let $G = (V,E)$ be a connected, undirected graph with Laplacian matrix $L = D - A$. Write $L = U \mu U^T$ for the spectral decomposition of $L$. Let the normalized Laplacian matrix be $N = D^{-1/2} L D^{-1/2}$, and write $N = V \Lambda V^T$ for the spectral decomposition of $N$. Let $M = D^{-1/2}ND^{1/2}$ be the random walk Laplacian (i.e. $M = I - W$) where $W$ is the random walk matrix.

\section{Graph clustering with \textrm{PPR}}

We begin with concrete tasks, specific to the editing of our paper ``Local Spectral Clustering of Density Upper Level Sets.''

\begin{enumerate}
	\item Rewrite introduction to carefully introduce and motivate our goals. Why, in light of what is already known about consistency of spectral clustering, consistency of density clustering in the flat and hierarchical setting, and spectral clustering in the nonparametric mixture models setting, is our study worthwhile?
	\item Relatedly, we should include a more extensive literature review to place our work in better context.
	\item Examples. To start, the Gaussian mixture and uniform mixture (with noise) case. 
\end{enumerate}

Equally concrete extensions, although not necessarily to be included in the aforementioned paper, include:

\begin{enumerate}
	\item Negative examples. Can we prove a lower bound on the symmetric set distance metric between \textrm{PPR}-recovered set and any high density set? Can we prove that the \textrm{PPR}-recovered set has higher normalized cut than a given high density set?
\end{enumerate}

Next, we detail some more abstract potential areas to study.

\subsection{Connection between spectral, local, and resistance-distance embeddings.}

We start by establishing a connection between spectral, random walk based (local), and resistance-distance based clustering approaches, based on embeddings associated with each approach.

\paragraph{Spectral Embedding.}
We start by recalling the definition of spectral clustering. Let $V_k = (v_1 \ldots v_k)$ be the $n \times k$ matrix containing the first $k$ eigenvectors of $N$. Letting $n = \abs{V}$, the spectral embedding into $\Reals^{k - 1}$ based on $V_k$ is given by $z_i = (v_{2i}, \ldots, v_{ki})$, so that
\begin{equation*}
\norm{z_i - z_j}_2^2 = (e_i - e_j)^T V_k V_k^T (e_i - e_j)
\end{equation*}
where $e_i$ is the $i$th standard basis vector in $\Reals^n$.

\paragraph{Resistance Distance Embedding.}
Let $N^{\dagger} = \sum_{k = 2}^{n} (1/\lambda_k) v_k v_k^T$ be the pseudoinverse of the Laplacian matrix. The resistance distance embedding into $\Reals^{n - 1}$ is given by $z_i = (\lambda_{2}^{-1/2}/\sqrt{d_i} u_{2i}, \ldots, \lambda_{n}^{-1/2}/\sqrt{d_i} u_{ni})$, so that
\begin{equation*}
\norm{z_i - z_j}_2^2 = (e_i - e_j)^T D^{-1/2} N^{\dagger} D^{-1/2} (e_i - e_j)
\end{equation*}
the latter being exactly the formula for resistance distance.

As shown by \citep{vonluxburg14}, the resistance distance embedding is not useful when $G$ is a neighborhood graph.

\paragraph{\textrm{PPR} resistance distance embedding.}

Given a parameter $\alpha > 0$ and seed node $e_i$, the \textrm{PPR} random walk is defined by the recursive relation:
\begin{equation*}
\textrm{pr}_{i}(\alpha) = \alpha e_i + (1 - \alpha) \textrm{pr}_{i}(\alpha) W
\end{equation*}

The following is a representation of the \textrm{PPR}~vector in terms of the inverse of the normalized Laplacian, from \citep{chung10}:
\begin{equation*}
\frac{(1 - \alpha)}{2\alpha}\textrm{pr}_{i}(\alpha) = e_i D^{-1/2}\left(N + \alpha I\right)^{-1} D^{1/2}
\end{equation*}
Following \citep{chung10}, we build to a notion of \textrm{PPR} resistance distance. We define the (normalized) \textrm{PPR} hitting time to be
\begin{equation*}
h_{\alpha}(i,j) = \frac{(\textrm{pr}_{j}(\alpha))_j}{d_j} -  \frac{(\textrm{pr}_{j}(\alpha))_i}{d_i}
\end{equation*}
and the \textrm{PPR} effective resistance to be
\begin{equation*}
R_{\alpha}(i,j) := h_{\alpha}(i,j) + h_{\alpha}(j,i).
\end{equation*}
It can be verified that
\begin{equation*}
R_{\alpha}(i,j) = \frac{(1- \alpha)}{2\alpha} (e_i - e_j)^T D^{-1/2} (N + \alpha I)^{-1} D^{-1/2} (e_i - e_j).
\end{equation*}

Therefore, the \textrm{PPR} resistance distance induces an embedding, $z_i = \bigl( (\lambda_2 + \alpha)^{-1/2}/\sqrt{d_i})v_{2i}, \ldots, (\lambda_n + \alpha)^{-1/2}/\sqrt{d_i})v_{ni} \bigr)$.

Several directions are now worth mentioning:
\begin{itemize}
	\item These relations hint at why clustering based on PPR or spectral embeddings may be useful, even though clustering based on the resistance distance is not. Both of the former two methods involve regularization over the eigenvalues, and in particular ensure that the top eigenvalue does not dominate. We could more formally study the properties of \textrm{PPR} resistance distance embedding -- framing it either via comparison to spectral embedding, or by contrasting it with (not \textrm{PPR}) resistance distance embeddings.
	\item This also is an interesting--if vague--connection between our work here and developments on graph testing. In both cases, shrinking the eigenvalues of $\Linv$ has proved essential to meaningful developments.
	\item This would also be a natural area to study if we decided to use embedding (synthesis) as an overarching theme.
\end{itemize}

\subsection{Effect of normalization of Laplacian on recovery of high-density clusters.}

We know (see \citep{vonluxburg07}) that the optimization problem
\begin{equation}
\label{eqn:laplacian_energy}
\min_{\mathbf{1}^T z = 0} z^T L z
\end{equation}
is a relaxation of the (NP-hard) problem
\begin{equation}
\label{eqn:ratio_cut}
\min_{\mathbf{1}^T z = 0, z_i \in \set{0,1}} z^T L z
\end{equation}
We term the solution to \eqref{eqn:ratio_cut} the minimum \textit{RatioCut}, noting that the minimization problem is equivalent to finding $\min_{A \subseteq V} \mathrm{RatioCut}(A)$ where $\mathrm{RatioCut}(A)$ is defined as
\begin{equation}
\mathrm{RatioCut}(A) := \mathrm{cut}(A,A^c)\left(\frac{1}{\abs{A}} + \frac{1}{\abs{A^c}}\right).
\end{equation}

Let $N = D^{-1/2}LD^{-1/2}$ be the normalized Laplacian. The optimization problem
\begin{equation}
\label{eqn:norm_laplacian_energy}
\min_{\mathbf{1}^T z = 0} z^T N z
\end{equation}
is a relaxation of the (NP-hard) problem
\begin{equation}
\label{eqn:normalized_cut}
\min_{\mathbf{1}^T z = 0, z_i \in \set{0,1}} z^T N z
\end{equation}
which is equivalent to finding $A \subseteq V$ which miniminizes $\mathrm{NCut}{A}$, for
\begin{equation*}
\mathrm{NCut}(A) := \mathrm{cut}(A,A^c)\left(\frac{1}{\vol(A)} + \frac{1}{\vol(A)}\right).
\end{equation*}

We immediately note that for two subsets $A$ and $A'$ which have equivalent \textrm{RatioCut}, whichever subset with higher average degree will have the lower \textrm{NCut}. In this sense, therefore, the problem \eqref{eqn:norm_laplacian_energy} seems to place ``more emphasis'' on finding a cluster with high average density than does the problem \eqref{eqn:laplacian_energy}. 
\begin{itemize}
	\item Can we extend this intuition to a formal mathematical statement?
	\item Is there a normalization of $L$ (for example, of the form $N^{\alpha} = D^{-\alpha}LD^{-\alpha}$) such that a problem equivalent to \eqref{eqn:norm_laplacian_energy} or \eqref{eqn:laplacian_energy} (or indeed \eqref{eqn:normalized_cut} or \eqref{eqn:ratio_cut}) recovers exactly a high density cluster?
	\item Can we make some practical recommendation about what the ``right'' normalization of the Laplacian is?
\end{itemize}

\subsection{Normalized cut recovery of the density cluster tree.}

In \cite{chaudhuri2010}, the authors state that analysis is made cleaner/easier by working with a hierarchical cluster tree object, rather than the flat density clustering problem. How can we make use of this in the study of the spectral properties of density clusters?

Let $C_f$ be the hierarchical cluster tree for a density function $f: [0,1] \to \Reals_{\geq 0}$ of a distribution $P$. Additionally, let $\eta_r: [0,1] \times [0,1]$ by the kernel function $\eta_r(x,y) = \mathbf{1}(\norm{x - y} \leq r)$.  For $A \subseteq [0,1]$, let the \emph{local normalized cut} functional be given by,
\begin{align*}
\mathrm{cut}_{r,\Pbb}(A) = \int_{A} \int_{A^c} \eta_r(x,y) \,dP(y) \,dP(x), & \quad \quad \mathrm{vol}_{r,\Pbb}(A) = \int_{A} \int_{[0,1]} \eta_r(x,y) \,dP(y) \,dP(x) \\ \mathrm{LCut}_{r,\Pbb}(A) & = \frac{\mathrm{cut}_{r,\Pbb}(A)}{\mathrm{vol}_{r,\Pbb}(A)}.
\end{align*}

Observe that if $C \subseteq C'$ are both objects in $C_f$, then for sufficiently small $r > 0$, $\mathrm{LCut}_{r,\Pbb}(C) \geq \mathrm{LCut}_{r,\Pbb}(C')$. In other words, the $\mathrm{LCut}$ functional respects the partial ordering induced by $C_f$ over sets $C \in C_f$.

\begin{itemize}
	\item How can we derive algorithmic consequences from this fact? To be concrete, can I design an algorithm which, given the normalized cut of every set $A \subseteq [0,1]$, can reconstruct the cluster tree $C_f$? If not, what additional information do I need?
	\item Clearly for $d \geq 1$, the \textrm{NCut} functional does not in general respect the partial ordering induced by $C_f$. Can we still do something; e.g. can we consistently estimate those relations in $C_f$ which are respected by the functional $\mathrm{LCut}_{r,\Pbb}(\cdot)$?
\end{itemize}

\section{Spectral Graph Testing.}

We observe $G = (V,E)$ with $V = [n]$, and in addition observe
\begin{equation*}
y_i = \beta_i + \sigma \varepsilon_i, \quad \varepsilon_i \sim \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1),
\end{equation*}

\begin{itemize}
	\item We have derived minimax optimal convergence rates, using a projection kernel onto eigenvectors of $L$, when $\beta = (\beta_1, \ldots, \beta_n)$ is in a discrete Sobolev class, assuming appropriate tail decay of the eigenvalues $\mu_i$. Can we extend this to the setting where $G$ is a random neighborhood graph, and $\beta = (f(x_1), \ldots, f(x_n))$ for $f$ in a continuous Sobolev class?
	\item We know that simply considering an IPM-type statistic of the form
	\begin{equation*}
	T = \sup_{\theta: \norm{B\theta}_2^2 \leq C_{n,r}^2} \dotp{\theta}{y}
	\end{equation*}
	will not lead to a minimax optimal test. When $C$ is chosen to contain $\beta$ with high probability, the function class $\set{\theta: \norm{B\theta}_2^2 \leq C^2}$ is too rich, and the variance of $T_C$ therefore too large.

	Consider instead the test statistic: for some $C > 0$
	\begin{equation*}
	T_C = \sup_{\theta: \norm{(B + C I)\theta}_2^2 \leq C_{n,r}^2} \dotp{\theta}{y}
	\end{equation*}
	which restricts the function class under consideration. 
	Can we show that this test statistic -- based on a regularization, rather than truncation, of the eigenvalues -- achieves similar results to the aforementioned projection kernel?
	\item In what practical cases does such a test statistic outperform other related test statistics? As some examples, \citep{ingster2009} proposes a test based on Fourier series projection kernel. \citep{ariascastro2018} analyzes the chi-squared test. \citep{schilling1986} proposes a two-sample test based on nearest neighbors, which can be easily converted into a goodness-of-fit test. \citep{gretton2012} analyzes the MMD in an appropriate Reproducing Kernel Hilbert Space. There are others. 
	\item Can we derive minimax optimal convergence rates when $\beta$ is in a discrete total variation class? This entails both proving upper bounds and matching lower bound. 
	\item Can we extend this to the random neighborhood graph setting, where $f$ is a bounded variation function?
\end{itemize}

\section{Embedding / Other topic to focus on here.}

TODO. (One note: maybe it's worth thinking about ways that our grid inequality can be of use?)

\bibliographystyle{plainnat}
\bibliography{future_directions_bibliography}

\end{document}

### 2019-08-02 notes/graph_testing.tex
updated weekly notes; began drafting master document

--- 2019-01-30 notes/graph_testing.tex
@@
 \usepackage{amsfonts, amsthm, amssymb}
-\usepackage{bm}
 \usepackage{graphicx}
@@
 \usepackage{enumerate}
+\usepackage[shortlabels]{enumitem}
+\usepackage{fullpage}
+\usepackage{mathtools}
 
@@
 
-\makeatletter
-\newcommand{\leqnomode}{\tagsleft@true}
-\newcommand{\reqnomode}{\tagsleft@false}
-\makeatother
+\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
 
@@
 \newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
-\newcommand{\Var}[1]{\mathrm{Var}\left( #1 \right)}
 \newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
@@
 \newcommand{\myexp}[1]{\exp \{ #1 \}}
-\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
-\newcommand{\dotp}[2]{\langle #1 , #2 \rangle}
 \newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
 \newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
-\newcommand{\defeq}{\overset{\mathrm{def}}{=}}
-\newcommand{\convweak}{\overset{w}{\rightharpoonup}}
-\newcommand{\dive}{\mathrm{div}}
-\newcommand{\Bin}{\mathrm{Bin}}
+\newcommand{\ext}[1]{\widetilde{#1}}
+\newcommand{\set}[1]{\left\{#1\right\}}
+\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
+\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
+\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
+\newcommand{\Var}{\mathrm{Var}}
+\newcommand{\Cov}{\mathrm{Cov}}
+\newcommand{\diam}{\mathrm{diam}}
 
@@
 \newcommand{\vol}{\text{vol}}
-\newcommand{\Bal}{\textrm{Bal}}
-\newcommand{\Cut}{\textrm{Cut}}
-\newcommand{\Ind}{\textrm{Ind}}
-\newcommand{\set}[1]{\left\{#1\right\}}
-\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
-\newcommand{\Perp}{\perp \! \! \! \perp}
-\newcommand{\Naturals}{\mathbb{N}}
+\newcommand{\spansp}{\mathrm{span}~}
+\newcommand{\1}{\mathbf{1}}
 
 \newcommand{\Linv}{L^{\dagger}}
-\newcommand{\tr}{\text{tr}}
-\newcommand{\h}{\textbf{h}}
-% \newcommand{\l}{\ell}
-\newcommand{\x}{\textbf{x}}
-\newcommand{\y}{\textbf{y}}
-\newcommand{\bl}{\bm{\ell}}
-\newcommand{\bnu}{\bm{\nu}}
-\newcommand{\Lx}{\mathcal{L}_X}
-\newcommand{\Ly}{\mathcal{L}_Y}
 \DeclareMathOperator*{\argmin}{argmin}
+\DeclareMathOperator*{\argmax}{argmax}
 
+\newcommand{\emF}{\mathbb{F}_n}
 \newcommand{\emG}{\mathbb{G}_n}
-\newcommand{\A}{\mathcal{A}}
+\newcommand{\emP}{\mathbb{P}_n}
 \newcommand{\F}{\mathcal{F}}
-\newcommand{\G}{\mathcal{G}}
-\newcommand{\X}{\mathcal{X}}
+\newcommand{\D}{\mathcal{D}}
+\newcommand{\R}{\mathcal{R}}
 \newcommand{\Rd}{\Reals^d}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\E}{\mathcal{E}}
 
-%%% Matrix related notation
-\newcommand{\Xbf}{\mathbf{X}}
-\newcommand{\Ybf}{\mathbf{Y}}
-\newcommand{\Zbf}{\mathbf{Z}}
-\newcommand{\Abf}{\mathbf{A}}
-\newcommand{\Dbf}{\mathbf{D}}
-\newcommand{\Wbf}{\mathbf{W}}
-\newcommand{\Lbf}{\mathbf{L}}
-\newcommand{\Ibf}{\mathbf{I}}
-\newcommand{\Bbf}{\mathbf{B}}
+%%% Vectors
+\newcommand{\thetast}{\theta^{\star}}
 
-%%% Vector related notation
-\newcommand{\lbf}{\bm{\ell}}
-\newcommand{\fbf}{\mathbf{f}}
+%%% Matrices
+\newcommand{\X}{X} % no bold
+\newcommand{\Y}{Y} % no bold
+\newcommand{\Z}{Z} % no bold
+\newcommand{\Lgrid}{L_{\grid}}
+\newcommand{\Dgrid}{D_{\grid}}
+\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}
 
-%%% Set related notation
-\newcommand{\Dset}{\mathcal{D}}
-\newcommand{\Aset}{\mathcal{A}}
-\newcommand{\Wset}{\mathcal{W}}
+%%% Sets and classes
+\newcommand{\Xset}{\mathcal{X}}
+\newcommand{\Sset}{\mathcal{S}}
+\newcommand{\Hclass}{\mathcal{H}}
+\newcommand{\Pclass}{\mathcal{P}}
 
-%%% Distribution related notation
+%%% Distributions and related quantities
 \newcommand{\Pbb}{\mathbb{P}}
+\newcommand{\Ebb}{\mathbb{E}}
 \newcommand{\Qbb}{\mathbb{Q}}
-% \newcommand{\Pr}{\mathrm{Pr}}}
 
-%%% Functionals
-\newcommand{\1}{\mathbf{1}}
+%%% Operators
+\newcommand{\Tadj}{T^{\star}}
+\newcommand{\dive}{\mathrm{div}}
+\newcommand{\dif}{\mathop{}\!\mathrm{d}}
+\newcommand{\gradient}{\mathcal{D}}
+\newcommand{\Hessian}{\mathcal{D}^2}
+
+%%% Misc
+\newcommand{\grid}{\mathrm{grid}}
+\newcommand{\critr}{R_n}
+\newcommand{\dx}{\,dx}
+\newcommand{\dy}{\,dy}
+\newcommand{\dr}{\,dr}
+\newcommand{\dxpr}{\,dx'}
+\newcommand{\dypr}{\,dy'}
+\newcommand{\wt}[1]{\widetilde{#1}}
+
+%%% Order of magnitude
+\newcommand{\soom}{\sim}
+
+% \newcommand{\span}{\textrm{span}}
+
 \newtheoremstyle{alden}
@@
 \theoremstyle{alden}
-\newtheorem{definition}{Definition}[section]
 
@@
 \newtheorem{assumption}{Assumption}
+\newtheorem{remark}{Remark}
+
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
 
 \theoremstyle{remark}
-\newtheorem{remark}{Remark}
 
 \begin{document}
-
-\title{\textcolor{red}{Testing with Graphs}}
+\title{Graph Testing}
 \author{Alden Green}
@@
 
-\paragraph{Two-sample non-parametric hypothesis-testing problem.}
+\section{Fixed graph testing.}
 
-For fixed integers $n_1 + n_2 = n$, let $\Xbf = \set{x_1, \ldots, x_{n_1}} \subset \Rd$ and $\Ybf = \set{y_1, \ldots, y_{n_2}}$ be sampled i.i.d from distributions $\Pbb$ and $\Qbb$ with density functions $p$ and $q$, respectively, both with support on $D \subset \Rd$.
-Our statistical problem is testing the null hypothesis $H_0: \Pbb = \Qbb$ vs.
-the alternative $H_1: \Pbb \neq \Qbb$, where our knowledge of $\Pbb$ and $\Qbb$ come from the samples $\Xbf$ and $\Ybf$.
-
-\paragraph{Integral Probability Metric.}
-Given $\F$ a class of real-valued bounded measurable functions on $D$, the \emph{integral probability metric} between $\Pbb$ and $\Qbb$ with respect to $\F$ is
+Suppose we observe $G = (V,E)$, an undirected graph over $V = [n]$.
+Let $B$ be the $m \times n$ incidence matrix of $G$, with singular value decomposition $B = U\Lambda^{1/2}V^T$, so that the Laplacian matrix $L = V \Lambda V^T$.
+For $i \in [n]$, we observe
 \begin{equation*}
-\gamma_{\F}(\Pbb,\Qbb) = \sup_{f \in \F}\abs{\int_{D} f d\Pbb - \int_{D} f d\Qbb}
+z_i = \beta_i + \varepsilon_i, \quad \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
+\end{equation*}
+Our statistical goal is hypothesis testing. We wish to distinguish
+\begin{equation*}
+\mathbf{H}_0: \norm{\beta}_2 = 0 \quad \textrm{vs.} \quad \mathbf{H}_a: \norm{\beta}_2 > 0.
+\end{equation*}
+We will evaluate our performance using the notion of \emph{worst-case error}. For a given ``function'' class $\mathcal{H}$, test function $\phi: \Reals^n \to \set{0,1}$, and $\epsilon > 0$, let
+\begin{equation*}
+\mathcal{R}_{\epsilon}(\phi;\mathcal{H})= \Ebb_{0}(\phi) + \sup_{\beta \in \mathcal{H}:\norm{\beta}_2 > \epsilon} \Ebb_{\beta}(1 - \phi)
 \end{equation*}
 
-One such IPM that has not received close attention until surprisingly recently is the \emph{(weighted) Sobolev IPM}.
-For $f \in L^2(\Dset)$ and $\rho$ a density function over $\Dset$, the ($\rho^2$-weighted) Sobolev 1,2 norm of $f$ is given by
+\subsection{Discrete Sobolev Classes.}
+
+The first ``function'' class we will consider will be balls in discrete Sobolev norms, with known smoothness. For $s,d$ positive integers, and radius $C_n > 0$, let
 \begin{equation*}
-\norm{f}_{1,2,\rho^2} := 
-\begin{cases}
-\int_{\Dset}\norm{\nabla_x f(x)}^2 \rho^2 dx,~ & \text{$f \in H^1(\Dset)$} \\
-\infty,~ & \text{$f \in L^2(\Dset) \setminus H^1(\Dset)$}
-\end{cases}
-\end{equation*}
-where $H^1(\Dset)$ is \textcolor{red}{the Sobolev space of $L^2(\Dset)$ functions with weak derivative $\nabla_x f(x) \in L^2(\Dset)$.}
-
-Consider the unit ball of $\norm{\cdot}_{1,2,\rho^2}$,
-\begin{equation*}
-\Wset^{1,2}(\Dset,\rho^2) := \set{f: \norm{f}_{1,2,\rho^2} \leq 1}.
-\end{equation*}
-The weighted Sobolev IPM is simply $\gamma_{\Wset^{1,2}(\Dset,\rho^2)}(\Pbb,\Qbb)$.
-
-\paragraph{Neighborhood graph.}
-
-Let $G_{n,r_n} = (V,E,w)$ denote a weighted, undirected graph constructed from the samples $\Zbf = \set{z_1, \ldots, z_n} = (\Xbf,\Ybf)$ where $V = \set{1, \ldots, n}$, and $w_{uv} = K(z_u, z_v):= k(\frac{\norm{z_u - z_v}}{\epsilon_n}) \geq 0$ for $u,v \in V$, and a particular kernel function $k$.
-Here $(u,v) \in E$ if and only if $w_{uv} > 0$.
-
-Motivated by the integral probability metric, our test statistics will be of the form
-\begin{equation*}
-\gamma_{\F_n}(\Pbb_n,\Qbb_n) = \sup_{f_n \in \F_n}\abs{\int_{D} f_n d\Pbb_n - \int_{D} f_n d\Qbb_n}
+\mathcal{S}^{s}_{d}(C_n) = \set{\beta: \norm{D^{(s)}\beta}_2 \leq C_n}
 \end{equation*}
@@
 \begin{equation*}
-\Pbb_n := \frac{1}{n_1}\sum_{i = 1}^{n_1} \delta_{x_i}, ~~ \Qbb_n := \frac{1}{n_2}\sum_{i = 1}^{n_2} \delta_{y_i}
+D^{(s)} = 
+\begin{cases}
+L^{s/2}, & \text{$s$ even} \\
+DL^{(s-1)/2}, & \text{$s$ odd}.
+\end{cases}
 \end{equation*}
-are the empirical distributions of $\Xbf$ and $\Ybf$, respectively, and $\F_n$ is a \textcolor{red}{class of functions $f_n: \Zbf \to \Reals$ exhibiting some regularity with respect to the neighborhood graph $G_{n,r}$.}
+We note that the constraint is equivalent to $\beta^T V \Lambda^s V \beta \leq C_n^2$.
+\subsection{Test statistic}
+Let $0 = \lambda_0 \leq \lambda_1 \leq \lambda_2 \leq \ldots \leq \lambda_{n - 1}$ denote the ordered eigenvalues of $L$, and let $v_k$ denote the eigenvector corresponding to $\lambda_k$.
+For a constant $C > 0$ to be specified later, let
+\begin{equation*}
+T_C = \sum_{k: \lambda_k^{s} \leq C^2} y_k^2
+\end{equation*}
+where
+\begin{equation*}
+y_k = \frac{1}{\sqrt{n}} \dotp{z}{v_k}.
+\end{equation*}
 
-\paragraph{Laplacian smoothing and Total Variation denoising.}
+We first compute the expectation $\Ebb(T_C)$. Let $\theta \in \Reals^n$ represent the expectation of $(y_k)$, meaning
+\begin{equation*}
+\theta_k = \frac{1}{\sqrt{n}}\dotp{\beta}{v_k}
+\end{equation*}
+and let $\Pi_C\theta$ have entries $(\Pi_C \theta)_k = \1(\lambda_k^{s} \leq C^2) \theta_k$. Finally, let $N(C) = \sharp\set{k: \lambda_k^{s} \leq C^2}$.
 
-For convenience, we number the edges $E  = (e_1, \ldots, e_m)$.
-We denote by $\Bbf \in \Reals^{m \times n}$ the edge incidence matrix of $G_{n,r}$, which for $k$th edge $e_k = (u,v)$ has $k$th row $\Bbf_{k} = (0,\ldots,-w_{uv},\ldots,w_{uv},\ldots,0)$ with a $-w_{uv}$ in the $u$th location, and a $w_{uv}$ in the $v$th location.
-The random (unnormalized) Laplacian matrix is then $\Lbf = \Bbf^T \Bbf$.
-We also introduce a \emph{label vector}, given by $\lbf = (\ell_1, \ldots, \ell_n)$ with
+\begin{lemma}
+	\label{lem:expectation}
+	For any $\beta \in \Reals^n$,
 	\begin{equation}
-\label{eqn: label_vector}
-\ell_k = 
-\begin{cases}
-\frac{n}{n_1},~~~ \text{$z_k \in \Xbf$} \\
--\frac{n}{n_2}, ~ \text{$z_k \in \Ybf$}
-\end{cases}
+	\label{eqn:expectation_1}
+	\Ebb(T_C) = \frac{N(C)}{n} + \norm{\Pi_C \theta}_2^2 
+	\end{equation}
+	If additionally $\beta \in \mathcal{H}$, the following lower bound holds:
+	\begin{equation}
+	\label{eqn:expectation_2}
+	\Ebb(T_C) \geq \frac{N(C)}{n} + \frac{\norm{\beta}^2}{n}  - \frac{C_n^2}{nC^{2}}
+	\end{equation}
+\end{lemma}
+\begin{proof}
+	We can write
+	\begin{align*}
+	\Ebb(T_C) & = \sum_{k: \lambda_k^s \leq C} \Ebb(y_k^2) \\
+	& = \frac{1}{n}\sum_{k: \lambda_k^s \leq C} \Ebb\bigl(\dotp{\beta}{v_k}^2 + \dotp{\varepsilon}{v_k}^2 + 2 \dotp{\varepsilon}{v_k} \dotp{\beta}{v_k}\bigr) \\
+	& = \sum_{k: \lambda_k^s \leq C}\theta_k^2 + \frac{1}{n} \\
+	& = \norm{\Pi_C \theta}^2 + \frac{N(C)}{n},
+	\end{align*}
+	showing \eqref{eqn:expectation_1}. Now, assuming, $\norm{D^{(s)}\beta}_2 \leq C_n$, we can further obtain
+	\begin{align*}
+	\norm{\Pi_C \theta}^2 & = \norm{\theta}^2 - \sum_{k:\lambda_k^s > C^2} \theta_k^2 \\
+	& \geq \norm{\theta}^2 - \frac{1}{C^{2}}\sum_{k:\lambda_k^s > C^2} \theta_k^2 \lambda_k^s \\
+	& \geq \norm{\theta}^2 -\frac{1}{nC^2} \beta^T V \Lambda^s V^T \beta \\
+	& \geq \norm{\theta}^2  - \frac{C_n^2}{nC^{2}}
+	\end{align*}
+	and \eqref{eqn:expectation_2} is shown.
+\end{proof}
+
+We now turn to computing the variance $\Var(T_C)$.
+\begin{lemma}
+	\label{lem:variance}
+	\begin{equation*}
+	\Var(T_C) = \frac{2N(C)}{n^2} + \frac{4 \norm{\Pi_C \theta}_2^2}{n}
+	\end{equation*}
+\end{lemma}
+\begin{proof}
+	To begin, we rewrite
+	\begin{align*}
+	T_C & = \sum_{k:\lambda_k^s \leq C^2} y_k^2 \\
+	& = \frac{1}{n} \sum_{k: \lambda_k^s \leq C^2} \dotp{z}{v_k}^2 \\
+	& = \frac{1}{n} \sum_{k: \lambda_k^s \leq C^2} z^T v_k v_k^T z \\
+	& =: \frac{1}{n} z^T P_{C} z.
+	\end{align*}
+	where $P_C := \sum_{k: \lambda_k^s \leq C^2} v_k v_k^T$. Therefore $\Var(T_C) = \Var(z^T P_{C} z)/n^2$. We expand $z = \beta + \varepsilon$ to obtain
+	\begin{align}
+	\Var(z^T P_{C} z) & = \Var((\beta + \epsilon)^T P_{C} (\beta + \epsilon)) \nonumber \\
+	& = \Var(\beta^T P_{C} \beta + 2 \epsilon P_{C} \beta + \epsilon^T P_{C} \epsilon) \nonumber \\
+	& = 4 \beta^T P_{C} I P_{C} \beta + \Var(\epsilon^T P_{C} \epsilon) + 4 \Cov(\epsilon P_{C} \beta, \epsilon^T P_{C} \epsilon) \nonumber \\
+	& = 4 n \norm{\Pi_C \theta}_2^2 + \Var(\epsilon^T P_{C} \epsilon) \label{eqn:variance_1}
+	\end{align}
+	where the last equality follows from the Gaussianity of $\epsilon$, as
+	\begin{equation*}
+	\Ebb\bigl( (\epsilon P_{C} \beta) (\epsilon^T P_{C} \epsilon) \bigr) = \sum_{i = 1}^{n} \sum_{j = 1}^{n} \sum_{k = 1}^{n} (P_C\beta)_k (P_C)_{ij} \Ebb(\epsilon_i \epsilon_j \epsilon_k) = 0.
+	\end{equation*}
+	Also by the Gaussianity of $\epsilon$, $\epsilon^T P_C \epsilon \sim \chi_{N(C)}^2$, and therefore $\Var(\epsilon^T P_{C} \epsilon) = 2N(C)$. Plugging back into \eqref{eqn:variance_1}, we obtain
+	\begin{equation*}
+	\Var(z^T P_{C} z) = 4 n \norm{\Pi_C \theta}_2^2 + 2N(C) 
+	\end{equation*}
+	and therefore the desired result is proved.
+\end{proof}
+
+We will consider now the test $\phi_C = \mathbf{1}\{T(C) \geq N(C)/n + \tau(b)\}$, where for $b \geq 1$, $\tau(b) = b \sqrt{2N(C)/n^2}$. We first upper bound the type I error.
+
+\begin{lemma}
+	\label{lem:type_I_error}
+	Under the null hypothesis $\beta = 0$, and for any $C > 0$,
+	\begin{equation*}
+	\Ebb_{\beta = 0}(\phi_C) \leq \frac{1}{b^2}. 
+	\end{equation*}
+\end{lemma}
+\begin{proof}
+	The desired result follows from Chebyshev's inequality,
+	\begin{align*}
+	\Ebb_{\beta = 0}(\phi) & = \Pbb_{\beta = 0}\bigl(T(C) \geq N(C)/n + \tau(b)\bigr) \\ 
+	& = \Pbb_{\beta = 0}\bigl(T(C) - \frac{N(C)}{n} \geq \tau(b)\bigr) \\
+	& \leq \Pbb_{\beta = 0}\bigl(\abs{T(C) - \frac{N(C)}{n}} \geq \tau(b)\bigr) \\
+	& \leq \frac{\Var_{\beta = 0}(T_C)}{\tau(b)^2} = \frac{1}{b^2}.
+	\end{align*}
+\end{proof}
+The calculation for the type II error will be slightly more involved.
+\begin{lemma}
+	\label{lem:type_II}
+	Let $b \geq 1$ be fixed. For every $\beta \in \mathcal{S}_d^s(C_n)$ such that
+	\begin{equation}
+	\label{eqn:type_II}
+	\frac{\norm{\beta}^2}{n} \geq 2b\sqrt{2\frac{N(C)}{n^2}} + \frac{C_n^2}{nC^2}
+	\end{equation}
+	we have that
+	\begin{equation*}
+	\Ebb_{\beta}(1 - \phi) \leq \frac{2}{b^2} + \frac{2}{b\sqrt{2N(C)}}.
+	\end{equation*}
+\end{lemma}
+\begin{proof}
+	Let $\Delta = \Ebb_{\beta}(T_C) - N(C)/n = \norm{\Pi_C \theta}^2$, and observe that by Lemma \ref{lem:expectation} and \eqref{eqn:type_II},
+	\begin{equation*}
+	\Delta \geq \frac{\norm{\beta}_2^2}{n} - \frac{C_n^2}{nC^2} \geq 2 \tau(b).
+	\end{equation*}
+	An application of Chebyshev's inequality yields
+	\begin{align*}
+	\Ebb_{\beta}\bigl(1 - \phi\bigr) & = \Pbb_{\beta}\bigl(T_C \leq N(C)/n + \tau(b)\bigr) \\
+	& = \Pbb_{\beta}\bigl(T_C - \Ebb_{\beta}(T_C) \leq \tau(b) - \Delta \bigr) \\
+	& \leq \Pbb_{\beta}\bigl(\abs{T_C - \Ebb_{\beta}(T_C)} \leq \Delta - \tau(b) \bigr) \tag{since $\Delta \geq \tau(b)$}	\\
+	& \leq \frac{\Var_{\beta}(T_C)}{(\Delta - \tau(b))^2} \\
+	& \leq 4\frac{\Var_{\beta}(T_C)}{\Delta^2} \tag{since $\Delta \geq 2\tau(b)$} \\
+	& \leq 4\frac{2N(C)/n^2 + \norm{\Pi_C\theta}_2^2/n}{\Delta^2}.
+	\end{align*}
+
+	We now handle each summand separately. For the first term, since $\Delta \geq 2 \tau(b)$, we have
+	\begin{equation*}
+	\frac{2N(C)}{n^2\Delta^2} \leq \frac{1}{2b^2}.
+	\end{equation*}
+
+	For the second term, since $\Delta = \norm{\Pi_C\theta}^2$, we have
+	\begin{align*}
+	\frac{\norm{\Pi_C\theta}_2^2/n}{\Delta^2} & \leq \frac{1}{n\Delta^2} \\
+	& \leq \frac{1}{2n\tau(b)} \\
+	& = \frac{1}{2b\sqrt{2N(C)}}.
+	\end{align*}
+\end{proof}
+
+To more explicitly specify the critical radius $\epsilon: \norm{\beta}_2 \geq \epsilon$, we will need to make an assumption on the relation between $N(C)$ and $C$. In particular, let $C = C^*$, where
+\begin{equation*}
+C^* = c\frac{(C_n n^{s/d})^{4s/(4s + d)}}{n^{s/d}}.
+\end{equation*}
+for some $c$ constant in $n$. We will assume the following bound on $N(C^*)$:
+\begin{enumerate}[label=(A\arabic*)]
+	\item Tail decay:
+	\label{asmp:tail_decay}
+	\begin{equation*}
+	N(C^*) \leq (C^*)^{d/s}n
+	\end{equation*}
+\end{enumerate}
+
+\begin{corollary}
+	\label{cor:critical_radius}
+	Fix $b \geq 1$, assume \ref{asmp:tail_decay}. Then, for any $C_n > 0$
+	\begin{equation*}
+	\sup_{\substack{\beta \in \mathcal{S}_d^{s}(C_n), \\ \norm{\beta}_2/n \geq \epsilon}} \Ebb_{\beta}(1 - \phi_{C^{\star}}) \leq \frac{2}{b^2} + \frac{2}{b\sqrt{N(C^{\star})}}
+	\end{equation*}
+	for every $\epsilon > 0$ such that 
+	\begin{equation*}
+	\epsilon^2 \geq (2b + 1) \left(\frac{(C_n n^{s/d})^{2d/(4s + d)}}{n}\right).
+	\end{equation*}
+\end{corollary}
+\begin{proof}
+	First, we verify that $\epsilon^2 \geq 2b\sqrt{2N(C^*)/n^2} + C_n^2/n(C^*)^2$. By Assumption \ref{asmp:tail_decay} and the choice of $C^{\star}$, we have
+	\begin{equation*}
+	N(C^*) \leq (C^*)^{d/s}n = (C_n n^{s/d})^{4d/(4s + d)}
+	\end{equation*}
+	and therefore
+	\begin{equation*}
+	2b\sqrt{\frac{N(C^*)}{n^2}} \leq \frac{2b}{n} (C_n n^{s/d})^{2d/(4s+d)}
+	\end{equation*}
+	Moving on to the second term, we have
+	\begin{align*}
+	\frac{C_n^2}{n(C^{\star})^2} & = \frac{C_n^2n^{2s/d}}{n(C_nn^{s/d})^{8s/(4s+d)}} \\
+	& = \frac{C_n^{2d/(4s + d)}n^{(s/d)2d/(4s + d)}}{n} \\
+	& = \frac{(C_n n^{s/d})^{2d/(4s + d)}}{n}
+	\end{align*}
+	and therefore $\epsilon \geq 2b\sqrt{2N(C^*)/n^2} + C_n^2/n(C^*)^2$. As a result, by Lemma \ref{lem:type_II} for any $\beta \in \mathcal{S}^s_d(C_n)$ such that $\norm{\beta}/n \geq \epsilon$, 
+	\begin{equation*}
+	\Ebb_{\beta}(1 - \phi_{C^{\star}}) \geq \frac{2}{b^2} + \frac{2}{b\sqrt{2N(C^{\star})}}.
+	\end{equation*}
+\end{proof}
+
+\section{Graph-based Tests over Function Spaces}
+
+Let $\mathcal{D} = [0,1]^d$. Suppose we observe the random design $x_1,\ldots,x_n$ independently sampled from probability measure $P$ over $[0,1]^d$. Additionally, for $i \in [n]$, we observe
+\begin{equation*}
+z_i = f(x_i) + \varepsilon_i, \quad \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1), \quad \varepsilon \perp x
+\end{equation*}
+where $f \in \mathcal{H} \subseteq L^2(\mathcal{D})$. Our statistical goal is hypothesis testing. We wish to distinguish:
+\begin{equation*}
+\mathbf{H}_0: \norm{f}_2 = 0 \quad \textrm{vs} \quad \mathbf{H}_{\textrm{a}}: \norm{f}_2 > 0. 
+\end{equation*}
+We will evaluate our performance using worst-case error: for a given function class $\mathcal{H}$, test function $\phi: \Reals^n \to \set{0,1}$ and $\epsilon> 0$, let
+\begin{equation*}
+\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \Ebb_{f = 0}(\phi) + \sup_{f \in \mathcal{H}: \norm{f}_2 > \epsilon} \Ebb_f(1 - \phi)
+\end{equation*}
+
+\subsection{Sobolev Spaces.}
+
+The first type of function class $\mathcal{H}$ we consider will be balls in Sobolev spaces.
+Let $s,d$ be known, fixed positive integers.
+For $f: \mathcal{D} \to \Reals$ locally summable, we use the multiindex notation $D^{\alpha}f$ to denote the $\alpha$th-weak partial derivative of $f$ (if one exists).
+Then, the Sobolev norm is
+\begin{equation*}
+\norm{f}_{W^{s,2}(\D)}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{D}} \norm{D^{\alpha}f}_2^2 \,dx
+\end{equation*}
+and the corresponding unit ball is $W^{s,2}(\D; L) = \set{f: \norm{f}_{W^{s,2}(\mathcal{D})} \leq L}$.
+
+\subsubsection{Projection-based test statistic.}
+
+For $x,y \in \mathcal{D}$ and radius $r > 0$ to be specified later, let $\eta_r(x,y) = \mathbf{1}(\norm{x - y}/ r \leq 1)$.
+and let $A$ be the $n \times n$ adjacency matrix with entries $A_{ij} = \eta_r(x_i,x_j)$.
+Let $B$ be the incidence matrix associated with $A$, and $L = B^TB$ be the corresponding Laplacian matrix.
+Write $B = U \Lambda^{1/2}V^T$ for the singular value decomposition of $B$.
+Then $L = V \Lambda V^T$ is the eigendecomposition of $L$ , where $\Lambda$ is a diagonal matrix of eigenvalues with diagonal entries $\lambda_1 \leq \lambda_2 \leq \ldots \leq \lambda_n$, and $V$ is an orthonormal matrix of eigenvectors.
+
+Our test statistic will be the norm of a projection of $z$ onto the subspace spanned by the first eigenvectors of $V$. In particular, for $C > 0$ to be specified later, our test statistic will be
+\begin{equation*}
+T_C := \sum_{k: \lambda_k^s \leq C^2} y_k^2, \quad y_k = \frac{1}{\sqrt{n}} z^T v_k.
+\end{equation*}
+For $b \geq 1$, let $\tau(b) = b\sqrt{2N(C)/n^2}$. Then our test will be $\phi_{C} = \mathbf{1}\set{T_C \leq N(C)/n + \tau(b)}$.
+
+In the following result, we prove that for an appropriate choice of threshold $C$ and radius $r$, the test $\phi_C$ achieves (nearly) minimax-optimal rates. Let
+\begin{equation*}
+C^{\star} = \frac{\left(n^{1 + s/d}r^{d/2 + s}\right)^{4s/(4s + d)}}{n^{s/d}}.
+\end{equation*}
+
+\begin{theorem}
+	\label{thm:sobolev_testing_rate}
+	Let $b \geq 1$ be a fixed constant, and let $s = 1$. For any sequence $r_n$ such that $(\log n/n)^{1/d}r \to \infty$, there exists a constant $c$ which depends only on $d$ and $L$ such that
+	\begin{equation}
+	\label{eqn:sobolev_testing_rate_1}
+	\mathcal{R}_{\epsilon}(\phi_{C^{\star}}; \mathcal{W}^{s,2}(\mathcal{D};L)) \leq c_1\left(\frac{1}{b^2} + \frac{1}{b\sqrt{N(C^{\star})}}\right) + o(1)
+	\end{equation}
+	for every $\epsilon$ such that
+	\begin{equation}
+	\label{eqn:sobolev_testing_rate}
+	\epsilon^2 \geq c_2 b \left(\frac{(b n^{1 + s/d}r^{d/2 + s})^{2d/(4s + d)}}{n}\right).
+	\end{equation}
+\end{theorem}
+
+\subsection{Proof of Theorem \ref{thm:sobolev_testing_rate}}
+Our goal is to upper bound
+\begin{equation*}
+\mathcal{R}_{\epsilon}(\phi_{C^{\star}}; W^{1,2}(\mathcal{D};L)) = \Ebb_{f = 0}(\phi_{C^{\star}}) + \sup_{f \in W^{1,2}(\mathcal{D};L): \norm{f}_2 > \epsilon} \Ebb_{f}(1 - \phi_{C^{\star}})
+\end{equation*}
+By Lemma \ref{lem:type_I_error}, we immediately have $\Ebb_{f = 0}(\phi_{C^{\star}}) \leq 1/b^2$, and we turn to the worst-case type II error.
+Fix any $f \in W^{1,2}(\mathcal{D};L)$ such that $\norm{f}_2 > \epsilon$, and let $\beta$ be the length-$n$ random vector with $j$th entry $\beta_j = f(x_j)$.
+Consider the following events:
+\begin{enumerate}[label=(E\arabic*)]
+	\item 
+	\label{event:discrete_sobolev_norm}
+	Discrete Sobolev norm of $\beta$: There exists a constant $c_1$ which depends only on $p_{\max}$ and $d$ such that
+	\begin{equation*}
+	\norm{B^{(s)}\beta}_2^2 \leq c_1 b^2 n^2 r^{d + 2s} 
+	\end{equation*}
+	\item 
+	\label{event:eigenvalue_tail_decay}
+	Eigenvalue tail decay:
+	\begin{equation*}
+	N(C^{\star}) \leq n(C^{\star})^{d/s}.
+	\end{equation*}
+	\item 
+	\label{event:l2_norm}
+	$L_2$ norm of $\beta$:
+	\begin{equation*}
+	\frac{\norm{\beta}_2^2}{n} \geq \frac{1}{2}\epsilon^2.
+	\end{equation*}
+\end{enumerate}
+
+Each of these events happen with high probability as $n \to \infty$.
+\begin{lemma}
+	\label{lem:high_probability_events}
+	With probability at least $1 - \frac{1}{b^2}$ as $n \to \infty$,
+	\begin{equation}
+	\label{eqn:discrete_sobolev_norm}
+	\norm{B\beta}_2^2 \leq c_1 b^2 n^2 r^{d + 2}.
 	\end{equation}
 
-Our test statistics $T_1(\lbf; G_{n,r})$ and $T_2(\lbf; G_{n,r})$ are defined as follows:
-\begin{align*}
-	T_1(\lbf; G_{n,r}) := \sup_{\fbf \in \Reals^n: \norm{\Bbf \fbf}_1 \leq C_{n,r}} ~ \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_k \\
-	T_2(\lbf; G_{n,r}) := \sup_{\fbf \in \Reals^n: \norm{\Bbf \fbf}_2^2 \leq C_{n,r}} ~ \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_k
-\end{align*}
-where \textcolor{red}{$C_{n,r} = $} and we write $\fbf = (f_1,\ldots, f_n)$. We note that, as promised, these satisfy the form
-\begin{align*}
-T_1(\lbf; G_{n,r}) & = \sup_{f_n \in TV_n}\abs{\int_{D} f_n d\Pbb_n - \int_{D} f_n d\Qbb_n} \\
-T_2(\lbf; G_{n,r}) & = \sup_{f_n \in \Wset_n}\abs{\int_{D} f_n d\Pbb_n - \int_{D} f_n d\Qbb_n}
-\end{align*}
-where $TV_n = \set{\fbf: \norm{\Bbf \fbf}_1 \leq C_{n,r}}$ and $\Wset_n = \set{\fbf: \norm{\Bbf \fbf}_2 \leq C_{n,r}}$.
+	With probability tending to one as $n \to \infty$,
+	\begin{equation}
+	\label{eqn:eigenvalue_tail_decay}
+	N(C^{\star}) \leq n(C^{\star})^{d/s}.
+	\end{equation}
 
-\section{Consistency under fixed alternative}
-
-\paragraph{Binomialized data model.}
-For technical reasons, we would like $z_1, \ldots, z_n$ to be independent and identically distributed. We consider the following generative model, which we term the \emph{binomialized data model}:
-
-Fix $n \in \Naturals > 0$, and $n_1 \sim \Bin(n,1/2),~ n_2 = n - n_1$.
-Then, let $x_1, \ldots, x_{n_1} \in \Rd$ be a sequence of i.i.d random points chosen according to $\Pbb$, and $y_1, \ldots, y_{n_2} \in \Rd$ a separate sequence of i.i.d random points chosen according to $\Qbb$, with $x_j \perp y_k$ for all $j,k$.
-Fix $\widetilde{\Zbf} = (\widetilde{z}_1,\ldots,\widetilde{z}_n) := (x_1, \ldots, x_{n_1}, y_1, \ldots, y_{n_2})$.
-Finally, for a permutation $\pi: [n] \to [n]$ chosen uniformly at random among all such permutations, let $\Zbf = (z_1,\ldots,z_n) =  (\widetilde{z}_{\pi(1)}, \ldots, \widetilde{z}_{\pi(n)})$.
-
-The label vector $\lbf$ remains defined as in \eqref{eqn: label_vector} with respect to $\Zbf$.
-Note that now $z_i \overset{i.i.d}{\sim} \frac{\Pbb}{2} + \frac{\Qbb}{2}$, as we desired, with density function $\mu(x) := \frac{p(x) + q(x)}{2}$.
-
-\begin{theorem}
-	\label{thm: LLN_for_t2}
-	Let $d \geq 2$ and let $\Dset \subset \Rd$ be an open, bounded, connected set with Lipschitz boundary. Let $\mu$ satisfy
+	\textcolor{red}{$L_2$ norm}.
+\end{lemma}
+In the following subsections, we prove Lemma \ref{lem:high_probability_events}.
+First, we show that Theorem \ref{thm:sobolev_testing_rate} is a simple consequence of Lemma \ref{lem:high_probability_events} along with our previous work.
+Let $\mathcal{A}_f$ be the set of $X$ such that \ref{event:discrete_sobolev_norm}, \ref{event:eigenvalue_tail_decay}, and \ref{event:l2_norm} all occur, for the given choice of $f$.
+We have that for $C_n := c_1bnr^{d/2 + s}$, the event \ref{event:discrete_sobolev_norm} implies $\beta \in \mathcal{S}_d^s(C_n)$.
+Therefore, by \ref{event:l2_norm},
+\begin{equation}
+\label{eqn:sobolev_testing_rate_2}
+\Ebb_f(1 - \phi_{C^*}) \leq \sup_{\substack{\beta \in \mathcal{S}_d^{s}(C_n), \\ \norm{\beta}_2/n \geq \epsilon^2/2}} \Ebb_{\beta}(1 - \phi_{C^{\star}}) + \Pbb(\mathcal{A}_f).
+\end{equation}
+Observe the following: 
+\begin{itemize}
+	\item \textcolor{red}{$C^{\star} = (C_n n^{s/d})^{4s/(4s+d)}/n^{s/d}$},
+	\item The event \ref{event:eigenvalue_tail_decay} is exactly \ref{asmp:tail_decay}, and 
+	\item For an appropriate choice of $c_2$, by hypothesis
 	\begin{equation*}
-	m \leq \mu(x) \leq M \tag{$\forall x \in D$}
+	\epsilon^2/2 \geq (2\sqrt{2}b + 1) \left(\frac{(C_n n^{s/d})^{2d/(4s + d)}}{n}\right)
 	\end{equation*}
-	for some $0 < m \leq M$. Let $(r_n)$ be a sequence of positive numbers converging to $0$ and satisfying
-	\begin{align*}
-	& \lim_{n \to \infty} \frac{(\log n)^{3/4}}{n^{1/2}} \frac{1}{r_n} = 0 ~~ \text{if $d = 2$} \\
-	& \lim_{n \to \infty} \frac{(\log n)^{1/d}}{n^{1/d}} \frac{1}{r_n} = 0 ~~ \text{if $d \geq 3$}
-	\end{align*}
-	Assume the kernel $k$ satisfies conditions: 
-	\begin{align*}
-	& k(0) > 0 ~ \text{and $k$ is continuous at $0$.} \tag{\textbf{K1}} \\
-	& k~ \text{is non-increasing.} \tag{\textbf{K2}} \\
-	& \text{The integral $\int_{0}^{\infty} k(r) r^{d+1} dr$ is finite.} \tag{\textbf{K3}}
-	\end{align*}
-
-	Then with probability one the following statement holds: For $(z_1, \ldots, z_n)$ chosen under the binomialized data model
+\end{itemize}
+Therefore we may apply Corollary \ref{cor:critical_radius}, and obtain
 \begin{equation*}
-	\lim_{n \to \infty} T_2(\lbf; G_{n,r_n}) = \gamma_{\Wset^{1,2}(\Dset,\mu^2)}(\Pbb,\Qbb).
+\sup_{\substack{\beta \in \mathcal{S}_d^{s}(C_n), \\ \norm{\beta}_2/n \geq \epsilon^2/2}} \Ebb_{\beta}(1 - \phi_{C^{\star}}) \leq \frac{2}{b^2} + \frac{2}{b\sqrt{N(C^{\star})}},
 \end{equation*}
-\end{theorem}
-
-\section{Proofs}
-
-Fix $\mu_n = \frac{d\Pbb_n + d\Qbb_n}{2}$. \textcolor{red}{We will show a variational form of convergence of $\mu_n$ to $\mu$.}
-
-\subsection{Gamma convergence of constraint}
-
-\begin{definition}[$TL^2$ convergence]
-	Denote by $\mathfrak{B}(\Dset)$ the Borel $\sigma$-algebra of $\Dset$ and $\mathcal{P}(\Dset)$ the set of all Borel probability measures on $\Dset$.
-Given a Borel map $T: \Dset \to \Dset$, the \emph{push-forward} of $\mu$ by $T$ is given by
+and along with \eqref{eqn:sobolev_testing_rate_2} and Lemma \ref{lem:high_probability_events}, this shows
 \begin{equation*}
-	T_{\star}\mu(\Aset) := \mu(T^{-1}(\Aset)), ~~ \Aset \in \mathfrak{B}(\Dset).
+\Ebb_f(1 - \phi_{C^*}) \leq \frac{3}{b^2} + \frac{2}{b\sqrt{N(C^{\star})}} + o(1).
 \end{equation*}
 
-	Given $\widetilde{\mu} \in \mathcal{P}(\Dset)$, we say that $T$ is a \emph{transportation map} between $\mu$ and $\widetilde{\mu}$ if $T_{\star}\mu = \widetilde{\mu}$.
-If for a sequence $(T_n)$ of transportation maps
-	\begin{equation*}
-	\int_{\Dset} \abs{x - T_n(x)}^2 d\mu(x) \to 0,~ \text{as $n \to \infty$}
-	\end{equation*}
-	we refer to the sequence as \emph{stagnating.}
-
-	Take $f \in L^2(\mu)$ and a sequence $(f_n)$ with $f_n \in L^2(\mu_n)$ for $n = 1,2,\ldots$. If there exists a stagnating sequence of transportation maps $(T_n)$ such that
-	\begin{equation*}
-	\int_{\Dset} \abs{ f - T_n(f_n(x)) }^2 d\mu(x) \to 0,~ \text{as $n \to \infty$}
-	\end{equation*}
-	we say that $(f_n)$ converges $TL^2$ to $f$, and write $f_n \overset{TL^2}{\to} f$.
-\end{definition}
-
-We restate Theorem 1.4 of \cite{trillos2018}, changing notation to match the rest of this paper.
-
-\begin{theorem}[Theorem 1.4 of \cite{trillos2018}]
-	Under the setup and conditions of Theorem \ref{thm: LLN_for_t2}, with probability one the following statements hold:
-	\begin{itemize}
-		\item \emph{Liminf inequality:}
-		For all $f \in L^2(\mu)$ and all sequences $(f_n)$ with $f_n \in L^2(\mu_n)$ and $f_n \overset{TL^2}{\to} f$,
-		\begin{equation*}
-		\liminf_{n \to \infty} \frac{1}{n^2 r_n^{d + 2}} \norm{Bf}_2^2 \geq \textcolor{red}{\sigma_{k}} \norm{f}_{1,2,\rho^2}
-		\end{equation*}
-
-		\item \emph{Limsup inequality:} For all $f \in L^2(\mu)$, there exists a sequence $(f_n)$ with $f_n \in L^2(\mu_n)$ and $f_n \overset{TL^2}{\to} f$ such that
-		\begin{equation*}
-		\limsup_{n \to \infty} \frac{1}{n^2 r_n^{d + 2}} \norm{Bf}_2^2 \leq \textcolor{red}{\sigma_{k}} \norm{f}_{1,2,\rho^2}
-		\end{equation*}
-
-		\item \emph{Compactness property:} Every sequence $(f_n)$ with $f_n \in L^2(\mu_n)$ satisfying
-		\begin{equation*}
-		\sup_{n \in \Naturals} \frac{1}{n^2 r_n^{d + 2}} \norm{Bf}_2^2 < \infty
-		\end{equation*}
-		is precompact in $TL^2$, that is, every subsequence of $(f_n)$ has a further subsequence which converges in the $TL^2$-sense to an element of $L^2(\Dset)$.
-	\end{itemize}
-
-\end{theorem}
-
-\subsection{Continuity of risk functional}
-
-\subsection{Proof of Theorem \ref{thm: LLN_for_t2}}
-
-\clearpage
-
-\bibliography{../graph_testing_bibliography}
-\bibliographystyle{plain}
+\subsection{Proof of \eqref{eqn:discrete_sobolev_norm}}
 

### 2019-08-02 notes/weekly_notes/7.17.19.NotesForWeek.tex
updated weekly notes; began drafting master document

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbb{I}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 7/17/19 - 7/31/19}
\author{Alden Green}
\date{\today}
\maketitle

\section{Goodness-of-Fit Testing in the Sampling Model.}

Let $\mathcal{D} = [0,1]^d$. Suppose we observe the random design $x_1,\ldots,x_n$ independently sampled from the uniform distribution over $[0,1]^d$. Additionally, for $i \in [n]$, we observe
\begin{equation*}
z_i = f(x_i) + \varepsilon_i, \quad \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1), \quad \varepsilon \perp x
\end{equation*}
where $f \in \mathcal{H} \subseteq L^2(\mathcal{D})$. Our statistical goal is hypothesis testing. We wish to distinguish:
\begin{equation*}
\mathbf{H}_0: \norm{f}_2 = 0 \quad \textrm{vs} \quad \mathbf{H}_a: \norm{f}_2 > 0. 
\end{equation*}
We will evaluate our performance using worst-case error: for a given function class $\mathcal{H}$, test function $\phi: \Reals^n \to \set{0,1}$ and $\epsilon> 0$, let
\begin{equation*}
\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \Ebb_{f = 0}(\phi) + \sup_{f \in \mathcal{H}: \norm{f}_2 > \epsilon} \Ebb_f(1 - \phi)
\end{equation*}

An example function class $\mathcal{H}$ we will consider will be the unit ball in a Sobolev space. Let $s,d$ be known, fixed positive integers. For $f: \mathcal{D} \to \Reals$ locally summable, we use the multiindex notation $D^{\alpha}f$ to denote the $\alpha$th-weak partial derivative of $f$ (if one exists). Then, the Sobolev norm is 
\begin{equation*}
\norm{f}_{W^{s,2}(\D)}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{D}} \norm{D^{\alpha}f}_2^2 \,dx
\end{equation*}
and the corresponding unit ball is $W^{s,2}(\D; L) = \set{f: \norm{f}_{W^{s,2}(\mathcal{D})} \leq L}$.

\subsection{Test statistic.}

For $x,y \in \mathcal{D}$ and radius $r > 0$ to be specified later, let $\eta_r(x,y) = \mathbf{1}(\norm{x - y}/ r \leq 1)$. and let $A$ be the $n \times n$ adjacency matrix with entires $A_{ij} = \eta_r(x_i,x_j)$. Let $B$ be the incidence matrix associated with $A$, and $L = B^TB$ be the associated Laplacian matrix. Write $B = U \Lambda^{1/2}V^T$ for the singular value decomposition of $B$. Then $L = V \Lambda V^T$ is the eigendecomposition of $L$ , where $\Lambda$ is a diagonal matrix of eigenvalues with diagonal entries $\lambda_1 \leq \lambda_2 \leq \ldots \leq \lambda_n$, and $V$ is an orthonormal matrix of eigenvectors.

Our test statistic will be the norm of a projection of $z$ onto the subspace spanned by the first few eigenvectors of $V$. In particular, for $C > 0$ to be specified later, our test statistic will be
\begin{equation*}
T_C := \sum_{k: \lambda_k^s \leq C^2} y_k^2, \quad y_k = \frac{1}{\sqrt{n}} z^T v_k.
\end{equation*}
For $b \geq 1$, let $\tau(b) = b\sqrt{2N(C)/n^2}$. Then our test will be $\phi_{C} = \mathbf{1}\set{T_C \leq N(C)/n + \tau(b)}$.

\subsection{Fixed design testing.}

Let $\beta$ be the length-$n$ random vector with $j$th entry $\beta_j = f(x_j)$. We introduce the scaling $C_{n,r} > 0$, defined by $C_{n,r}^2 = n^2r^{d + 2s}$.  We will apply results about testing over balls in discrete Sobolev classes. Let $b \geq 1$ denote $N(C) := \sharp\set{k: \lambda_k^s \leq C^2}$, and consider the following events:
\begin{enumerate}[label=(E\arabic*)]
	\item 
	\label{event:discrete_sobolev_norm}
	Discrete Sobolev norm of $\beta$:
	\begin{equation*}
	\norm{B^{(s)}\beta}_2 \leq C_{n,r}
	\end{equation*}
	\item Eigenvalue tail decay:
	\label{event:tail_decay}
	For the choice $C = C_r^{\star}$, where
	\begin{equation*}
	C_r^{\star} = \frac{(C_{n,r}n^{s/d})^{4s/(4s + d)}}{n^{s/d}},
	\end{equation*}
	the following inequality is satisfied:
	\begin{equation*}
	N(C_r^{\star}) \leq n(C_r^{\star})^{d/s}.
	\end{equation*}
	\item 
	\label{event:l2_norm}
	$L_2$ norm of $\beta$:
	\begin{equation*}
	\frac{\norm{\beta}_2^2}{n} \geq \frac{(2\sqrt{2}b + 1)}{n} \left(C_{n,r}n^{s/d}\right)^{2d/(4s+d)}.
	\end{equation*}
\end{enumerate}

Suppose for a given $x = x_1, \ldots, x_n$, and for a particular choice of radius $r$, the events \ref{event:l2_norm}-\ref{event:tail_decay} hold. The worst-case error conditional on $x$ can then be upper bounded.
\begin{lemma}
	\label{lem:discrete_test_error}
	For any $x$ and $r$ such that \ref{event:discrete_sobolev_norm} and \ref{event:tail_decay} hold, we have that if $f = 0$,
	\begin{equation*}
	\Ebb_{\beta}(\phi|x) \leq \frac{1}{b^2}
	\end{equation*}
	If $f \neq 0$ is such that \ref{event:l2_norm} is additionally satisfied, we have that
	\begin{equation*}
	\Ebb_{\beta}(1 -\phi|x) \leq \frac{1}{2b^2} + o(1)
	\end{equation*}
\end{lemma}

We now turn to showing that each of \ref{event:discrete_sobolev_norm} and \ref{event:tail_decay} (and, under appropriate conditions on $\norm{f}_2$, \ref{event:l2_norm}) hold with probability at least $1 - \frac{1}{b^2} - o(1)$.

\section{Bounding discrete Sobolev norm.}
Recall that our goal is to show \ref{event:discrete_sobolev_norm} occurs with high probability. We will build slowly to this goal.

\subsection{$s = 1$, $f$ Lipschitz.}
To start, we provide a bound in the case when $s = 1$, and $f$ has bounded Lipschitz norm. Precisely, we define the space of $1$-Holder (Lipschitz) functions $C_{\mathcal{X}}^{0,1}(1)$ to consist of all continuous functions $g: \D \to \Reals$ such that
\begin{equation*}
\norm{g}_{C_{\D}^{0,1}} := \norm{g}_{C(\D)} + [g]_{C_{\D}^{0,1}} \leq 1
\end{equation*}
where $\norm{g}_{C(\D)} = \sup_{x \in \D} \abs{g(x)}$, and
\begin{equation*}
[g]_{C_{\D}^{0,1}} = \sup_{x,y \in \D} \frac{\abs{g(x) - g(y)}}{\norm{x - y}}.
\end{equation*}

Let $P_n$ be the empirical distribution associated with $x_1,\ldots,x_n$, i.e. $P_n = \frac{1}{n}\sum_{i = 1}^{n} \delta_{x_i}$. We begin by providing a deterministic bound involving the distance between the measures $P_n$ and $P$. As these are measures over different spaces, it is not obvious how to relate them. We will use transportation distance to do so. Recall that a mapping $T: \D \to \D$ is a \emph{transportation map} between $P$ and $P_n$ if for all measurable $A \subseteq \D$,
\begin{equation*}
P_n(A) = P(T^{-1}(A))
\end{equation*}

\begin{lemma}
	\label{lem:holder_1_bound}
	For any $f \in C_{\mathcal{D}}^{0,1}$, the following bound holds on the Sobolev discrete norm:
	\begin{equation*}
	\norm{B\beta}_2^2 \leq n^2r^2(r + \norm{\mathrm{Id} - T}_{L^{\infty}(P)})^d
	\end{equation*}
\end{lemma}
\begin{proof}
	We write
	\begin{align*}
	\frac{1}{n^2}\norm{B^{(s)}\beta}_2^2 & = \frac{1}{n^2} \beta^T L \beta \\
	& = \frac{1}{2n^2} \sum_{i,j = 1}^{n} (\beta_i - \beta_j)^2 A_{ij} \\
	& = \int_{\D} \int_{\D} (f(x) - f(y))^2 \eta_r(x,y) \,dP_n(y) \,dP_n(x)
	\end{align*}
	We examine the inner integral. By the Holder property of $f$, and a change of variables, we obtain
	\begin{align*}
	\int_{\D} (f(x) - f(y))^2 \eta_r(x,y) \,dP_n(y) & \leq \int_{\D} \norm{x - y}^2 \eta_r(x,y) \,dP_n(y) \\
	& = \int_{\D} \norm{x - T(y)}^2 \eta_r(x,T(y)) \,dP(y) \\
	& \leq r^2 \int_{\D} \eta_r(x,T(y)) \,dP(y) \\
	& \leq r^2 \bigl( r + \norm{\mathrm{Id} - T}_{L^{\infty}(P)} \bigr)^d
	\end{align*}
	and the desired result is shown.
\end{proof}

\subsection{$s = 1$, $f \in \mathcal{W}^{1,2}(\D; 1)$.}

Let $P$ be an absolutely continuous probability measure over $\D$ with density function $p$ bounded above and below by constants, i.e
\begin{equation*}
0 < p_{\textrm{min}} < p(x) < p_{\textrm{max}} < \infty, \quad \textrm{for all $x \in \mathcal{D}$}
\end{equation*}

\begin{lemma}
	\label{lem:sobolev_1_bound}
	For any $f \in W^{1,2}(\mathcal{D};1)$, and any $b \geq 1$, we have that there exists a constant $C > 0$ which depends only on $d$ and $p_{\max}$ such that
	\begin{equation}
	\label{eqn:sobolev_1_bound}
	\norm{B\beta}_2^2 \leq b C n^2 r^{d + 2}
	\end{equation}
	with probability at least $1 - \frac{1}{b}$. 
\end{lemma}
\begin{proof}
	Observe that
	\begin{align*}
	\frac{1}{n^2}\Ebb(\beta^T L \beta) = \Ebb\left(\frac{1}{n^2} \sum_{i,j = 1}^{n} (\beta_i - \beta_j)^2 A_{ij}\right) & = \frac{(n - 1)}{n} \int_{\D} \int_{\D} (f(x) - f(y))^2\eta_r(x,y) \,dP(y) \,dP(x) \\
	& \leq p_{\max}^2 \frac{(n - 1)}{n} \int_{\D} \int_{\D} (f(x) - f(y))^2\eta_r(x,y) \,dy \,dx.
	\end{align*}
	We will show that for any $f \in \mathcal{W}^{1,2}(\D;1)$, there exists a constant $C$ which depends only on dimension $d$ such that
	\begin{equation}
	\label{eqn:mean_bound}
	\int_{\D} \int_{\D} (f(x) - f(y))^2\eta_r(x,y) \,dy \,dx \leq C r^{d + 2}
	\end{equation}
	whence the desired result of \eqref{eqn:sobolev_1_bound} follows by Markov's inequality.

	We begin by dealing with complications due to the boundary of $\mathcal{D}$. Let $V$ be any bounded open set such that $\D \subset \subset V$. Note that as $\partial \D$ is $C^1$, by Theorem \ref{thm:evans_extension} there exists $g \in W^{1,2}(\Reals^d)$ such that
	\begin{enumerate}
		\item
		\label{eqn:evans_extension_1}
		$g = f$, $P$-almost-everywhere in $\D$
		\item 
		$g$ has support within $V$, and  
		\item 
		\label{eqn:sobolev_1_bound_2}
		$\norm{g}_{W^{1,2}(\Rd)} \leq C \norm{f}_{W^{1,2}(\D)}$ for a constant $C$ which depends only on $\mathcal{D}$.
	\end{enumerate}
	As a result of the first point,
	\begin{equation}
	\label{eqn:sobolev_1_bound_1}
	\int_{\D} \int_{\D} (f(x) - f(y))^2\eta_r(x,y) \,dy \,dx \leq \int_{\Rd} \int_{\Rd} (g(x) - g(y))^2\eta_r(x,y) \,dy \,dx.
	\end{equation}

	Next, we smooth $g$, so that we may work with ordinary partial derivatives.
	We let $\kappa \in C^{\infty}(\Rd)$ be given by
	\begin{equation*}
	\kappa(x) :=
	\begin{cases}
	C \exp \left\{\frac{1}{\norm{x}^2 - 1}\right\}& \quad \textrm{if $\norm{x}_2 \leq 1$} \\
	0 & \quad \textrm{if $\norm{x}_2 \geq 1$}
	\end{cases}
	\end{equation*}
	where the normalizing constant $C > 0$ is chosen so that $\int_{\Rd} \eta dx = 1$. Let $\kappa_r(x) := (1/r^d) \kappa(x/r)$. Then, the mollification of $g$ by $\kappa_r$ is given by
	\begin{align*}
	g^r & := g \ast \eta_r \\
	& = \int_{\Rd} \eta_r(x - y) g(y) dy
	\end{align*}
	(Refer to \citep{evans10}, Appendix C, Theorem 7 for a proof that $g^r \in C^{\infty}(\Rd)$.)
	Adding and subtracting within \eqref{eqn:sobolev_1_bound_1}, we have
	\begin{align}
	\int_{\Rd} & \int_{\Rd} (g(x) - g(y))^2\eta_r(x,y) \,dy \,dx \\
	& \leq \int_{\Rd} \int_{\Rd} \bigl((g(x) - g^r(x))^2 + (g^r(x) - g^r(y))^2 + (g^r(y) - g(y))^2\bigr)\eta_r(x,y) \,dy \,dx \nonumber \\
	& = 2 \int_{\Rd} \int_{\Rd} \bigl((g(y) - g^r(y))^2 \eta_r(x,y) \,dy \,dx + \int_{\Rd} \int_{\Rd} \bigl((g^r(x) - g^r(y))^2 \eta_r(x,y) \,dy \,dx \label{eqn:sobolev_1_bound_5}
	\end{align}
	We deal with each summand individually, beginning with the first one. We have
	\begin{align}
	\int_{\Rd} \int_{\Rd} \bigl((g(y) - g^r(y))^2 \eta_r(x,y) \,dy \,dx & = \int_{\Rd} \int_{B(x,r)} \bigl((g(y) - g^r(y))^2\,dy \,dx \nonumber \\
	& = \int_{\Rd} \int_{B(y,r)} \bigl((g(y) - g^r(y))^2\,dx \,dy \tag{Tonelli's Theorem} \nonumber \\
	& = r^d \int_{\Rd}\bigl((g(y) - g^r(y))^2 \,dy \\
	& \leq r^{d + 2} \int_{\Rd} \norm{\nabla g(y)}^2 \,dy \label{eqn:sobolev_1_bound_4}
	\end{align}
	where the last line follows from Lemma \ref{lem:poincare_mollify}.

	We now turn our attention to the second summand. Note that as $g^r \in C^{\infty}(\Rd)$, we may apply Theorem \ref{thm:taylor_expansion} and obtain
	\begin{align*}
	(g^r(x) - g^r(y))^2 & = \left(\int_{0}^{1} \nabla g^r(x + t(y - x)) \cdot (y - x) \,dt\right)^2 \\
	& \leq \int_{0}^{1} \bigl(\nabla g^r(x + t(y - x)) \cdot (y - x)\bigr)^2 \,dt  \tag{Jensen's inequality} \\
	& \leq \norm{y - x}^2 \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 \,dt. \tag{Cauchy-Schwarz inequality}
	\end{align*}
	As a result, we have
	\begin{align*}
	\int_{\Rd} \int_{\Rd} \bigl((g^r(x) - g^r(y))^2 \eta_r(x,y) \,dy) \,dx & \leq \int_{\Rd} \int_{\Rd} \norm{y - x}^2  \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 \eta_r(x,y) \,dt \,dy \,dx \\
	& \leq r^{2} \int_{\Rd} \int_{\Rd} \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 \eta_r(x,y) \,dt \,dy \,dx \\
	& = r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + t(y - x))}^2 \eta_r(x,y) \,dx \,dt \,dy \\
	& = r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + tz)}^2 \eta_r(z) \,dx \,dt \,dz \tag{$z = y - x$}
	\end{align*}

	where we write $\eta_r(z) = \mathbf{1}(\norm{z} \leq r)$ in an abuse of notation. Next, we note that
	\begin{equation*}
	\int_{\Rd} \norm{\nabla g^r(x + tz)}^2 \,dx = \int_{\Rd} \norm{\nabla g^r(x)}^2 \,dx \leq \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
	\end{equation*}
	with the inequality following from Lemma \ref{lem:gradient_mollify_commute}. Therefore,
	\begin{align}
	r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + tz)}^2 \eta_r(z) \,dx \,dt \,dz & \leq r^{2} \int_{\Rd} \eta_r(z) \int_{0}^{1} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx  \,dt \,dz \nonumber \\
	& = r^{2 + d} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx. \label{eqn:sobolev_1_bound_3}
	\end{align}
	By \eqref{eqn:sobolev_1_bound_5}, \eqref{eqn:sobolev_1_bound_4} and \eqref{eqn:sobolev_1_bound_3}, we have that 
	\begin{equation*}
	\int_{\Rd} \int_{\Rd} (g(x) - g(y))^2 \eta_r(x,y) \,dx \,dy \leq 3 \int_{\Rd} \norm{\nabla g(x)}_2^2 \,dx
	\end{equation*}
	Then by \eqref{eqn:sobolev_1_bound_2}, $\int_{\Rd} \norm{\nabla g(x)}_2^2 \,dx \leq C\int_{\D} \norm{\nabla f(x)}_2^2 \,dx \leq C$ where $C$ is a constant depending only on $\D$. So the desired result of \eqref{eqn:sobolev_1_bound_1} follows.
\end{proof}

\section{Supporting Results.}

\begin{theorem}[\citep{evans10} Chapter 5.4, Theorem 1]
	\label{thm:evans_extension}
	Assume $U$ is bounded and $\partial U$ is $C^1$. Select a bounded open set $V$ such that $U \subset \subset V$ ($U$ is compactly contained in $V$). Then there exists a bounded linear operator $E: W^{1,2}(U) \to W^{1,2}(\Rd)$ such that for each $u \in W^{1,2}(U)$:
	\begin{enumerate}
		\item $Eu = u$ a.e. in $U$,
		\item $Eu$ has support within $V$, and 
		\item 
		\begin{equation*}
		\norm{Eu}_{W^{1,2}(\Rd)} \leq C \norm{u}_{W^{1,2}(\Rd)}
		\end{equation*}
		the constant $C$ depending only on $U$ and $V$.
	\end{enumerate}
\end{theorem}

For $u \in W^{1,2}(\Rd)$ and $x \in \Rd$, write $\nabla u(x) = (D^{e_1}(x),\ldots,D^{e_d}(x)$, for the gradient of $u$.

\begin{theorem}[\citep{evans10} Chapter 5.8.1, Theorem 2]
	\label{thm:evans_poincare}
	There exists a constant $C$, depending only on $d$, such that
	\begin{equation*}
	\norm{u - u^r}_{L^2(B(x,r))} \leq Cr\norm{\nabla u}_{L^2(B(x,r))}
	\end{equation*}
\end{theorem}

\begin{theorem}[Taylor expansion.]
	\label{thm:taylor_expansion}
	For any function $u \in C^1$, and any $x,y \in \Rd$, 
	\begin{equation*}
	u(y) - u(x) = \int_{0}^{1} \nabla(u(x + t(y - x))) \cdot (y - x) \,dt
	\end{equation*}
\end{theorem}

\begin{lemma}
	\label{lem:poincare_mollify}
	For any function $g \in W^{1,2}(\Rd)$ compactly supported in a bounded open set $V \subset \Rd$, we have
	\begin{equation}
	\label{eqn:poincare_mollify}
	\int_{\Rd} (g(x) - g^r(x))^2 \,dx \leq r^2 \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
	\end{equation}
\end{lemma}
\begin{proof}
	This Lemma is essentially a reproduction of part of the proof of the Rellich-Kondrachov Compactness Theorem from \citep{evans10}. Note that it is sufficient to prove in the case when $g$ is smooth. To see this, for the moment assume \eqref{eqn:poincare_mollify} holds for all $u \in C^{\infty}(V)$, and let $g \in W^{1,2}(V)$. By Theorem \ref{thm:local_approx_smooth_functions}, we may take a sequence $g_m \in C^{\infty}(V)$ such that
	\begin{equation*}
	\norm{g - g_m}_{L^2(V)} \to 0, \quad \textrm{and} \quad \norm{\nabla g - \nabla g_m}_{L^2(V)} \to 0,
	\end{equation*}
	Then we have
	\begin{align*}
	\int_{\Rd} (g(x) - g^r(x))^2 \,dx & \leq \int_{\Rd} (g(x) - g_m(x))^2 \,dx + \int_{\Rd}(g_m(x) - g_m^r(x))^2 \,dx + \int_{\Rd}(g_m^r(x) - g^r(x))^2 \,dx \\
	& \leq \int_{\Rd} (g(x) - g_m(x))^2 \,dx + r^2 \int_{\Rd} \norm{\nabla g_m(x)}^2 \,dx + \int_{\Rd}(g_m^r(x) - g^r(x))^2 \,dx
	\end{align*}
	and taking the limit as $m$ goes to infinity, the right hand side converges to $\int_{\Rd} \norm{\nabla g(x)}^2 \,dx$.

	It remains to show \eqref{eqn:poincare_mollify} in the case where $g$ is smooth. In this case, we have
	\begin{align*}
	g^{r}(x) -  g(x) & = \frac{1}{r^d} \int_{B(x,r)} \kappa \left(\frac{x - z}{r}\right)\bigl(g(z) - g(x)\bigr) \,dz \\
	& = \int_{B(0,1)} \kappa(y) \bigl(g(x - ry) - g(x)\bigr) \,dy \\
	& = \int_{B(0,1)} \kappa(y) \int_{0}^{1} \frac{d}{dt} \bigl(g(x - try) \bigr) \,dt \,dy \\
	& = -r \int_{B(0,1)} \kappa(y) \int_{0}^{1} \bigl(\nabla g (x - try) \bigr) \cdot y \,dt \,dy.
	\end{align*} 
	Therefore, by Jensen's and Cauchy-Schwarz inequalities, we have
	\begin{align*}
	\int_{\Rd} (g(x) - g^r(x))^2 \,dx & \leq r^2 \int_{\Rd} \int_{B(0,1)} \kappa(y) \int_{0}^{1} \norm{\nabla g (x - try)}^2 \norm{y}^2 \,dt \,dy \,dx \\
	& \leq r^2 \int_{B(0,1)} \kappa(y) \int_{0}^{1} \int_{\Rd} \norm{\nabla g (x - try)}^2 \,dx \,dt \,dy \\
	& = r^2 \int_{\Rd} \norm{\nabla g(z)}^2 \,dz
	\end{align*}
\end{proof}

The following theorem is Theorem 1 in Section 5.3 of \citep{evans10}.
\begin{theorem}[Local approximation by smooth functions.]
	\label{thm:local_approx_smooth_functions}
	Assume $U$ is bounded, and $u \in W^{k,p}(U)$ for some $1 \leq p < \infty$. Then there exists functions $u_m \in C^{\infty}(U) \cap W^{k,p}(U)$ such that
	\begin{equation*}
	\norm{u_m - u}_{W^{k,p}(U)} \overset{m}{\to} 0.
	\end{equation*}
\end{theorem}

\begin{lemma}
	\label{lem:gradient_mollify_commute}
	For any $u \in W^{1,2}(\Rd)$, we have
	\begin{equation*}
	\int_{\Reals^d} \norm{\nabla u^r(x)}_2^2 \,dx \leq \int_{\Reals^d} \norm{\nabla u(x)}_2^2 \,dx
	\end{equation*}
\end{lemma}
\begin{proof}
	Observe that $\norm{\nabla u(x)}_2^2 = \sum_{j = 1}^{d} (D^{e_j}u(x))^2 $. Therefore,
	\begin{align*}
	\int_{\Rd} \norm{\nabla u^r(x)}_2^2 \,dP(x) & = \sum_{j = 1}^{d} \int_{\Rd} (D^{e_j}u^r(x))^2 \,dx \\
	& = \sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u)^r(x))^2 \,dx
	\end{align*}
	where the second equality follows from equation (1) in Section 5.3 of \citep{evans10}. Then, for any $v \in L^2(\Rd)$, we have that
	\begin{align*}
	\abs{v^r(x)} & = \int_{\Rd} \kappa_r^{1/2}(x - y) \kappa_r^{1/2}(x - y) v(y) \,dy \\
	& \leq \left(\int_{\Rd} \kappa_r^(x - y) \,dy\right)^{1/2} \left(\int_{\Rd} \kappa_r(x - y) v^2(y) \right)^{1/2} \\
	& = \left(\int_{\Rd} \kappa_r(x - y) v^2(y) \right)^{1/2}
	\end{align*}
	and therefore
	\begin{align*}
	\int_{\Rd} (v^r(x))^2 \,dx & \leq \int_{\Rd} \int_{\Rd } \kappa_r(x - y) v^2(y) \,dy \,dx \\
	& = \int_{\Rd} \int_{\Rd } \kappa_r(x - y) v^2(y) \,dy \,dx \\
	& = \int_{\Rd} v^2(y) \,dy
	\end{align*}

	Applying this to $D^{e_j}u \in L^2(\Rd)$, we have that
	\begin{equation*}
	\sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u)^r(x))^2 \,dx \leq \sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u(x))^2 \,dx = \int_{\Rd} \norm{\nabla u(x)}^2 \,dx.
	\end{equation*}
\end{proof}

\bibliographystyle{plainnat}
\bibliography{../../graph_testing_bibliography}

\end{document}

### 2019-08-02 notes/weekly_notes/8.1.19.NotesForWeek.tex
updated weekly notes; began drafting master document

--- 2019-01-30 notes/graph_testing.tex
@@
 \newtheorem{assumption}{Assumption}
+\newtheorem{remark}{Remark}
+
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
 
 \theoremstyle{remark}
-\newtheorem{remark}{Remark}
 
 \begin{document}
-
-\title{\textcolor{red}{Testing with Graphs}}
+\title{Notes for Week 8/1/19 - 8/8/19}
 \author{Alden Green}
@@
 
-\paragraph{Two-sample non-parametric hypothesis-testing problem.}
+\section{Bound on Eigenvalue Tail Decay of Neighborhood Graph}
+Let $\mathcal{D} = [0,1]^d$. We consider two graphs over data on $\mathcal{D}$.
 
-For fixed integers $n_1 + n_2 = n$, let $\Xbf = \set{x_1, \ldots, x_{n_1}} \subset \Rd$ and $\Ybf = \set{y_1, \ldots, y_{n_2}}$ be sampled i.i.d from distributions $\Pbb$ and $\Qbb$ with density functions $p$ and $q$, respectively, both with support on $D \subset \Rd$.
-Our statistical problem is testing the null hypothesis $H_0: \Pbb = \Qbb$ vs.
-the alternative $H_1: \Pbb \neq \Qbb$, where our knowledge of $\Pbb$ and $\Qbb$ come from the samples $\Xbf$ and $\Ybf$.
+Suppose we observe the random design $X = x_1, \ldots, x_n$ independently sampled from probability measure $P$ supported on $\mathcal{D}$.
+For any $r \geq 0$, let the kernel function $\eta_r: \mathcal{D} \times \mathcal{D} \to \Reals_{\geq 0}$ be given by $\eta_r(x,y) = \1(\norm{x - y} \leq r)$.
+Then, let the \emph{r-neighborhood graph} over $x_1,\ldots,x_n$ be the undirected, unweighted graph $G = (V,E)$, where $V = [n]$ and for $i,j \in [n]$, $(i,j) \in E$ if $\eta_r(x_i, x_j) = 1$.
 
-\paragraph{Integral Probability Metric.}
-Given $\F$ a class of real-valued bounded measurable functions on $D$, the \emph{integral probability metric} between $\Pbb$ and $\Qbb$ with respect to $\F$ is
+Let $B$ be the incidence matrix associated with $G$, and let $L = B^TB$ be the corresponding Laplacian.
+Write $L = V \Lambda V^T$ for the eigen decomposition of $L$, where $V = (v_1 \ldots v_n)$ is an orthonormal matrix with the eigenvectors of $L$ as its columns, and $\Lambda$ is a diagonal matrix with entries $\lambda_1 \leq \lambda_2 \leq \ldots \leq \lambda_n$.
+For a given $s > 0$ and $C > 0$, let $N(C) = \#\set{k: \lambda_k^s \leq C^2}$.
+We wish to prove the following result.
+
+\begin{theorem}
+	\label{thm:spectral_decay}
+	Let $r \to 0$ as $n \to \infty$ sufficiently slowly so that $r(\log n/n)^{1/d} \to \infty$. Then, for each $s > 0$ and $C \leq \sqrt{2}$,
+	\begin{equation}
+	\label{eqn:spectral_decay}
+	N(C) \leq n C^{d/s}.
+	\end{equation} 
+	with probability tending to one as $n \to \infty$. 
+\end{theorem}
+
+\section{Theory}
+To prove Theorem \ref{thm:spectral_decay}, we will make use of another graph on points within $\mathcal{D}$, whose spectral properties are very well-understood.
+Let $\xi$ be the set of evenly spaced lattice points over $\mathcal{D}$; formally $\xi = \set{k/n: k \in [\ell]^d}$ where $\ell = n^{1/d}$, and we define $[\ell]^d = \set{(\ell_1,\ldots,\ell_d): \ell_k \in [\ell] ~\textrm{for each}~ \ell_k}$.
+Then, let the grid graph over $\xi$ be given by $\wt{G} = (\wt{V},\wt{E})$, where $\wt{V} = \xi$ and $(\xi_k, \xi_{k'})$ is in $\wt{E}$ if $\norm{\xi_k - \xi_{k'}}_1 = 1/\ell$.
+
+Let $\wt{L}$ be the Laplacian of $\wt{G}$, and let $\wt{\lambda_1} \leq \ldots \leq \wt{\lambda}_n$ be the ordered eigenvalues of $\wt{L}$.
+Write $\wt{N}(C) = \#\set{k: \wt{\lambda}_k^s \leq C^2}$.
+We have that the desired scaling rate \eqref{eqn:spectral_decay} holds with respect to the grid graph $\wt{G}$.
+
+\begin{lemma}
+	\label{lem:spectral_decay_grid}
+	For each $s$ and every $C \leq \sqrt{2}$, we have
+	\begin{equation}
+	\label{eqn:spectral_decay_grid}
+	\wt{N}(C) \leq 2^d\big(n C^{d/s} + 1)
+	\end{equation}
+\end{lemma}
+\begin{proof}
+	It will be sufficient to show that 
+	\begin{equation}
+	\label{eqn:spectral_decay_grid_1}
+	\lambda_k^s \leq C^2 \Rightarrow \floor{k^{1/d}}^d \leq C^{d/s}n
+	\end{equation}
+	To show this, observe that for any $\tau \in \mathbb{N}$ and $k = \tau^d$, we have
 	\begin{equation*}
-\gamma_{\F}(\Pbb,\Qbb) = \sup_{f \in \F}\abs{\int_{D} f d\Pbb - \int_{D} f d\Qbb}
+	\lambda_k \geq 4 \sin^2 \left(\frac{\pi k^{1/d}}{2n^{1/d}}\right) \geq \frac{\pi^2 k^{2/d}}{4 n^{2/d}} \wedge 2
+	\end{equation*}
+	Therefore, if $\lambda_k^s \leq C^2$ and $C \leq \sqrt{2}$, this implies
+	\begin{equation*}
+	\frac{\pi^{2s} k^{2s/d}}{4^s n^{2s/d}} \leq C^{2}
+	\end{equation*}
+	and rearranging, we obtain
+	\begin{equation*}
+	k \leq \frac{C^{d/s} 2^d n}{\pi^d} \leq C^{d/s}n
+	\end{equation*}	
+	If $k^{1/d}$ is not a natural number, applying the same argument to $k' = \floor{k^{1/d}}^d$ yields \eqref{eqn:spectral_decay_grid_1}.
+\end{proof}
+
+In light of Lemma \ref{lem:spectral_decay_grid}, to prove Theorem \ref{thm:spectral_decay} it is sufficient to show that $\wt{L} \preceq L$, since by the Courant-Fischer min-max theorem, the ordering $\wt{L} \preceq L$ implies that $\wt{\lambda}_k \leq \lambda_k$ for all $k \in [n]$.
+The next result details the conditions under which this ordering holds.
+This condition will be stated with respect to the min-max matching distance between $\xi$ and $X$, i.e.
+the minimum over all bijections $T: \xi \to X$ such that
+\begin{equation*}
+\min_{T} \max_{i \in [n]} \abs{T^{-1}(x_i) - x_i} 
 \end{equation*}
 
-One such IPM that has not received close attention until surprisingly recently is the \emph{(weighted) Sobolev IPM}.
-For $f \in L^2(\Dset)$ and $\rho$ a density function over $\Dset$, the ($\rho^2$-weighted) Sobolev 1,2 norm of $f$ is given by
+\begin{lemma}
+	\label{lem:partial_ordering_grid}
+	For any radius $r$ satisfying
+	\begin{equation}
+	\label{eqn:radius_condition}
+	r \geq 2 \min_{T} \max_{i \in [n]} \abs{T^{-1}(x_i) - x_i} + n^{-1/d}
+	\end{equation}
+	we have that $\wt{L} \preceq L$. 
+\end{lemma}
+\begin{proof}
+	Let $T_{\star}$ achieve the min-max matching distance, i.e
 	\begin{equation*}
-\norm{f}_{1,2,\rho^2} := 
-\begin{cases}
-\int_{\Dset}\norm{\nabla_x f(x)}^2 \rho^2 dx,~ & \text{$f \in H^1(\Dset)$} \\
-\infty,~ & \text{$f \in L^2(\Dset) \setminus H^1(\Dset)$}
-\end{cases}
-\end{equation*}
-where $H^1(\Dset)$ is \textcolor{red}{the Sobolev space of $L^2(\Dset)$ functions with weak derivative $\nabla_x f(x) \in L^2(\Dset)$.}
-
-Consider the unit ball of $\norm{\cdot}_{1,2,\rho^2}$,
-\begin{equation*}
-\Wset^{1,2}(\Dset,\rho^2) := \set{f: \norm{f}_{1,2,\rho^2} \leq 1}.
-\end{equation*}
-The weighted Sobolev IPM is simply $\gamma_{\Wset^{1,2}(\Dset,\rho^2)}(\Pbb,\Qbb)$.
-
-\paragraph{Neighborhood graph.}
-
-Let $G_{n,r_n} = (V,E,w)$ denote a weighted, undirected graph constructed from the samples $\Zbf = \set{z_1, \ldots, z_n} = (\Xbf,\Ybf)$ where $V = \set{1, \ldots, n}$, and $w_{uv} = K(z_u, z_v):= k(\frac{\norm{z_u - z_v}}{\epsilon_n}) \geq 0$ for $u,v \in V$, and a particular kernel function $k$.
-Here $(u,v) \in E$ if and only if $w_{uv} > 0$.
-
-Motivated by the integral probability metric, our test statistics will be of the form
-\begin{equation*}
-\gamma_{\F_n}(\Pbb_n,\Qbb_n) = \sup_{f_n \in \F_n}\abs{\int_{D} f_n d\Pbb_n - \int_{D} f_n d\Qbb_n}
-\end{equation*}
-where
-\begin{equation*}
-\Pbb_n := \frac{1}{n_1}\sum_{i = 1}^{n_1} \delta_{x_i}, ~~ \Qbb_n := \frac{1}{n_2}\sum_{i = 1}^{n_2} \delta_{y_i}
-\end{equation*}
-are the empirical distributions of $\Xbf$ and $\Ybf$, respectively, and $\F_n$ is a \textcolor{red}{class of functions $f_n: \Zbf \to \Reals$ exhibiting some regularity with respect to the neighborhood graph $G_{n,r}$.}
-
-\paragraph{Laplacian smoothing and Total Variation denoising.}
-
-For convenience, we number the edges $E  = (e_1, \ldots, e_m)$.
-We denote by $\Bbf \in \Reals^{m \times n}$ the edge incidence matrix of $G_{n,r}$, which for $k$th edge $e_k = (u,v)$ has $k$th row $\Bbf_{k} = (0,\ldots,-w_{uv},\ldots,w_{uv},\ldots,0)$ with a $-w_{uv}$ in the $u$th location, and a $w_{uv}$ in the $v$th location.
-The random (unnormalized) Laplacian matrix is then $\Lbf = \Bbf^T \Bbf$.
-We also introduce a \emph{label vector}, given by $\lbf = (\ell_1, \ldots, \ell_n)$ with
-\begin{equation}
-\label{eqn: label_vector}
-\ell_k = 
-\begin{cases}
-\frac{n}{n_1},~~~ \text{$z_k \in \Xbf$} \\
--\frac{n}{n_2}, ~ \text{$z_k \in \Ybf$}
-\end{cases}
-\end{equation}
-
-Our test statistics $T_1(\lbf; G_{n,r})$ and $T_2(\lbf; G_{n,r})$ are defined as follows:
-\begin{align*}
-	T_1(\lbf; G_{n,r}) := \sup_{\fbf \in \Reals^n: \norm{\Bbf \fbf}_1 \leq C_{n,r}} ~ \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_k \\
-	T_2(\lbf; G_{n,r}) := \sup_{\fbf \in \Reals^n: \norm{\Bbf \fbf}_2^2 \leq C_{n,r}} ~ \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_k
-\end{align*}
-where \textcolor{red}{$C_{n,r} = $} and we write $\fbf = (f_1,\ldots, f_n)$. We note that, as promised, these satisfy the form
-\begin{align*}
-T_1(\lbf; G_{n,r}) & = \sup_{f_n \in TV_n}\abs{\int_{D} f_n d\Pbb_n - \int_{D} f_n d\Qbb_n} \\
-T_2(\lbf; G_{n,r}) & = \sup_{f_n \in \Wset_n}\abs{\int_{D} f_n d\Pbb_n - \int_{D} f_n d\Qbb_n}
-\end{align*}
-where $TV_n = \set{\fbf: \norm{\Bbf \fbf}_1 \leq C_{n,r}}$ and $\Wset_n = \set{\fbf: \norm{\Bbf \fbf}_2 \leq C_{n,r}}$.
-
-\section{Consistency under fixed alternative}
-
-\paragraph{Binomialized data model.}
-For technical reasons, we would like $z_1, \ldots, z_n$ to be independent and identically distributed. We consider the following generative model, which we term the \emph{binomialized data model}:
-
-Fix $n \in \Naturals > 0$, and $n_1 \sim \Bin(n,1/2),~ n_2 = n - n_1$.
-Then, let $x_1, \ldots, x_{n_1} \in \Rd$ be a sequence of i.i.d random points chosen according to $\Pbb$, and $y_1, \ldots, y_{n_2} \in \Rd$ a separate sequence of i.i.d random points chosen according to $\Qbb$, with $x_j \perp y_k$ for all $j,k$.
-Fix $\widetilde{\Zbf} = (\widetilde{z}_1,\ldots,\widetilde{z}_n) := (x_1, \ldots, x_{n_1}, y_1, \ldots, y_{n_2})$.
-Finally, for a permutation $\pi: [n] \to [n]$ chosen uniformly at random among all such permutations, let $\Zbf = (z_1,\ldots,z_n) =  (\widetilde{z}_{\pi(1)}, \ldots, \widetilde{z}_{\pi(n)})$.
-
-The label vector $\lbf$ remains defined as in \eqref{eqn: label_vector} with respect to $\Zbf$.
-Note that now $z_i \overset{i.i.d}{\sim} \frac{\Pbb}{2} + \frac{\Qbb}{2}$, as we desired, with density function $\mu(x) := \frac{p(x) + q(x)}{2}$.
-
-\begin{theorem}
-	\label{thm: LLN_for_t2}
-	Let $d \geq 2$ and let $\Dset \subset \Rd$ be an open, bounded, connected set with Lipschitz boundary. Let $\mu$ satisfy
-	\begin{equation*}
-	m \leq \mu(x) \leq M \tag{$\forall x \in D$}
-	\end{equation*}
-	for some $0 < m \leq M$. Let $(r_n)$ be a sequence of positive numbers converging to $0$ and satisfying
-	\begin{align*}
-	& \lim_{n \to \infty} \frac{(\log n)^{3/4}}{n^{1/2}} \frac{1}{r_n} = 0 ~~ \text{if $d = 2$} \\
-	& \lim_{n \to \infty} \frac{(\log n)^{1/d}}{n^{1/d}} \frac{1}{r_n} = 0 ~~ \text{if $d \geq 3$}
-	\end{align*}
-	Assume the kernel $k$ satisfies conditions: 
-	\begin{align*}
-	& k(0) > 0 ~ \text{and $k$ is continuous at $0$.} \tag{\textbf{K1}} \\
-	& k~ \text{is non-increasing.} \tag{\textbf{K2}} \\
-	& \text{The integral $\int_{0}^{\infty} k(r) r^{d+1} dr$ is finite.} \tag{\textbf{K3}}
-	\end{align*}
-
-	Then with probability one the following statement holds: For $(z_1, \ldots, z_n)$ chosen under the binomialized data model
-	\begin{equation*}
-	\lim_{n \to \infty} T_2(\lbf; G_{n,r_n}) = \gamma_{\Wset^{1,2}(\Dset,\mu^2)}(\Pbb,\Qbb).
-	\end{equation*}
-\end{theorem}
-
-\section{Proofs}
-
-Fix $\mu_n = \frac{d\Pbb_n + d\Qbb_n}{2}$. \textcolor{red}{We will show a variational form of convergence of $\mu_n$ to $\mu$.}
-
-\subsection{Gamma convergence of constraint}
-
-\begin{definition}[$TL^2$ convergence]
-	Denote by $\mathfrak{B}(\Dset)$ the Borel $\sigma$-algebra of $\Dset$ and $\mathcal{P}(\Dset)$ the set of all Borel probability measures on $\Dset$.
-Given a Borel map $T: \Dset \to \Dset$, the \emph{push-forward} of $\mu$ by $T$ is given by
-	\begin{equation*}
-	T_{\star}\mu(\Aset) := \mu(T^{-1}(\Aset)), ~~ \Aset \in \mathfrak{B}(\Dset).
+	\max_{i \in [n]} \abs{T_{\star}^{-1}(x_i) - x_i} = \min_{T} \max_{i \in [n]} \abs{T^{-1}(x_i) - x_i}.
 	\end{equation*}
 
-	Given $\widetilde{\mu} \in \mathcal{P}(\Dset)$, we say that $T$ is a \emph{transportation map} between $\mu$ and $\widetilde{\mu}$ if $T_{\star}\mu = \widetilde{\mu}$.
-If for a sequence $(T_n)$ of transportation maps
+	It will be sufficient to prove that for every pair $(T_{\star}^{-1}(x_i), T_{\star}^{-1}(x_j)) \in \wt{E}$, the corresponding edge $(i,j)$ is in $E$.
+To see this, let $A$ denote the adjacency matrix associated with the neighborhood graph $G$, and $\wt{A}$ the adjacency matrix associated with the grid $\wt{G}$.
+Precisely $A$ is the $n \times n$ matrix with entries $A_{ij} = \eta_r(x_i,x_j)$, and $\wt{A}$ is the $n \times n$ matrix with entries $\wt{A}_{ij} = \1\{T_{\star}^{-1}(x_i), T_{\star}^{-1}(x_j) \in \wt{E} \}$.
+Our goal is to show that, for every $z \in \Rd$, we have
 	\begin{equation*}
-	\int_{\Dset} \abs{x - T_n(x)}^2 d\mu(x) \to 0,~ \text{as $n \to \infty$}
+	z^T L z = \frac{1}{2}\sum_{i, j = 1}^{n} (z_i - z_j)^2 A_{ij} \leq \frac{1}{2}\sum_{i, j = 1}^{n} (z_i - z_j)^2 \wt{A}_{ij} = z^T \wt{L} z.
 	\end{equation*}
-	we refer to the sequence as \emph{stagnating.}
+	which certainly holds if $\wt{A}_{ij} \leq A_{ij}$ for all $i,j \in [n]$.
 
-	Take $f \in L^2(\mu)$ and a sequence $(f_n)$ with $f_n \in L^2(\mu_n)$ for $n = 1,2,\ldots$. If there exists a stagnating sequence of transportation maps $(T_n)$ such that
+	Now, assume $(T_{\star}^{-1}(x_i), T_{\star}^{-1}(x_j)) \in \wt{E}$. This implies
+	\begin{align*}
+	\norm{x_i - x_j}_2 & \leq \norm{x_i - T_{\star}^{-1}(x_i)}_2 + \norm{T_{\star}^{-1}(x_j) - T_{\star}^{-1}(x_j)}_2 + \norm{x_j - T_{\star}^{-1}(x_j)}_2 \\
+	& \leq 2 \max_{i \in [n]} \abs{T_{\star}^{-1}(x_i) - x_i} + n^{-1/d} \\
+	& \leq r.
+	\end{align*}
+	so $\eta_r(x_i,x_j) = 1$ and therefore $(i,j) \in E$. 
+\end{proof}
+
+Finally, the following Lemma demonstrates that, for sufficiently large $r$, the condition \eqref{eqn:radius_condition} will hold with high probability.
+
+\begin{lemma}
+	\label{lem:matching_distance}
+	Assume $P$ has density $p$ which is bounded above and below uniformly over $\D$; that is, there exist constants $p_{\min}$ and $p_{\max}$ such that
 	\begin{equation*}
-	\int_{\Dset} \abs{ f - T_n(f_n(x)) }^2 d\mu(x) \to 0,~ \text{as $n \to \infty$}
+	0 < p_{\min} < p(x) < p_{\max} < \infty, \quad \text{for all $x \in \mathcal{D}$}.
 	\end{equation*}
-	we say that $(f_n)$ converges $TL^2$ to $f$, and write $f_n \overset{TL^2}{\to} f$.
-\end{definition}
+	Then, for any $r = r_n$ such that $r(\log(n)/n)^{1/d} \to \infty$, we have that the event
+	\begin{equation*}
+	r < 2 \min_{T} \max_{i \in [n]} \abs{T^{-1}(x_i) - x_i} + n^{-1/d}
+	\end{equation*}
+	occurs with probability tending to $0$ as $n \to \infty$. 
+\end{lemma}
 
-We restate Theorem 1.4 of \cite{trillos2018}, changing notation to match the rest of this paper.
-
-\begin{theorem}[Theorem 1.4 of \cite{trillos2018}]
-	Under the setup and conditions of Theorem \ref{thm: LLN_for_t2}, with probability one the following statements hold:
-	\begin{itemize}
-		\item \emph{Liminf inequality:}
-		For all $f \in L^2(\mu)$ and all sequences $(f_n)$ with $f_n \in L^2(\mu_n)$ and $f_n \overset{TL^2}{\to} f$,
-		\begin{equation*}
-		\liminf_{n \to \infty} \frac{1}{n^2 r_n^{d + 2}} \norm{Bf}_2^2 \geq \textcolor{red}{\sigma_{k}} \norm{f}_{1,2,\rho^2}
-		\end{equation*}
-
-		\item \emph{Limsup inequality:} For all $f \in L^2(\mu)$, there exists a sequence $(f_n)$ with $f_n \in L^2(\mu_n)$ and $f_n \overset{TL^2}{\to} f$ such that
-		\begin{equation*}
-		\limsup_{n \to \infty} \frac{1}{n^2 r_n^{d + 2}} \norm{Bf}_2^2 \leq \textcolor{red}{\sigma_{k}} \norm{f}_{1,2,\rho^2}
-		\end{equation*}
-
-		\item \emph{Compactness property:} Every sequence $(f_n)$ with $f_n \in L^2(\mu_n)$ satisfying
-		\begin{equation*}
-		\sup_{n \in \Naturals} \frac{1}{n^2 r_n^{d + 2}} \norm{Bf}_2^2 < \infty
-		\end{equation*}
-		is precompact in $TL^2$, that is, every subsequence of $(f_n)$ has a further subsequence which converges in the $TL^2$-sense to an element of $L^2(\Dset)$.
-	\end{itemize}
-
-\end{theorem}
-
-\subsection{Continuity of risk functional}
-
-\subsection{Proof of Theorem \ref{thm: LLN_for_t2}}
-
-\clearpage
-
-\bibliography{../graph_testing_bibliography}
-\bibliographystyle{plain}
+Together Lemmas \ref{lem:spectral_decay_grid}, \ref{lem:partial_ordering_grid}, and \ref{lem:matching_distance} imply Theorem \ref{thm:spectral_decay}.
 

### 2019-08-05 notes/graph_testing.tex
updated graph testing document

@@
 \section{Fixed graph testing.}
+\label{sec:fixed_graph_testing}
 
@@
 \begin{equation*}
+\label{eqn:C_star}
 C^* = c\frac{(C_n n^{s/d})^{4s/(4s + d)}}{n^{s/d}}.
@@
 	\begin{equation*}
-	\sup_{\substack{\beta \in \mathcal{S}_d^{s}(C_n), \\ \norm{\beta}_2/n \geq \epsilon}} \Ebb_{\beta}(1 - \phi_{C^{\star}}) \leq \frac{2}{b^2} + \frac{2}{b\sqrt{N(C^{\star})}}
+	\sup_{\substack{\beta \in \mathcal{S}_d^{s}(C_n), \\ \norm{\beta}_2/\sqrt{n} \geq \epsilon}} \Ebb_{\beta}(1 - \phi_{C^{\star}}) \leq \frac{2}{b^2} + \frac{2}{b\sqrt{N(C^{\star})}}
 	\end{equation*}
@@
 	\begin{equation*}
-	\epsilon^2 \geq (2b + 1) \left(\frac{(C_n n^{s/d})^{2d/(4s + d)}}{n}\right).
+	\epsilon^2 \geq (2bc^{d/2s} + c^{-2}) \left(\frac{(C_n n^{s/d})^{2d/(4s + d)}}{n}\right),
 	\end{equation*}
+	where $c$ is the same constant as appears in \eqref{eqn:C_star}.
 \end{corollary}
 \begin{proof}
-	First, we verify that $\epsilon^2 \geq 2b\sqrt{2N(C^*)/n^2} + C_n^2/n(C^*)^2$. By Assumption \ref{asmp:tail_decay} and the choice of $C^{\star}$, we have
+	Our goal is to verify that, when
 	\begin{equation*}
-	N(C^*) \leq (C^*)^{d/s}n = (C_n n^{s/d})^{4d/(4s + d)}
+	\frac{\norm{\beta}_2^2}{n} \geq (2bc^{d/2s} + c^{-2}) \left(\frac{(C_n n^{s/d})^{2d/(4s + d)}}{n}\right)
+	\end{equation*} 
+	then \eqref{eqn:type_II} holds with respect to $C^{\star}$, so that we may apply Lemma \ref{lem:type_II}.
+In other words, we want to show $\epsilon^2 \geq 2b\sqrt{2N(C^*)/n^2} + C_n^2/n(C^*)^2$.
+By Assumption \ref{asmp:tail_decay} and the choice of $C^{\star}$, we have
+	\begin{equation*}
+	N(C^*) \leq (C^*)^{d/s}n = c^{d/s}(C_n n^{s/d})^{4d/(4s + d)}
 	\end{equation*}
@@
 	\begin{equation*}
-	2b\sqrt{\frac{N(C^*)}{n^2}} \leq \frac{2b}{n} (C_n n^{s/d})^{2d/(4s+d)}
+	2b\sqrt{\frac{N(C^*)}{n^2}} \leq \frac{2bc^{d/2s}}{n} (C_n n^{s/d})^{2d/(4s+d)}
 	\end{equation*}
@@
 	\begin{align*}
-	\frac{C_n^2}{n(C^{\star})^2} & = \frac{C_n^2n^{2s/d}}{n(C_nn^{s/d})^{8s/(4s+d)}} \\
-	& = \frac{C_n^{2d/(4s + d)}n^{(s/d)2d/(4s + d)}}{n} \\
-	& = \frac{(C_n n^{s/d})^{2d/(4s + d)}}{n}
+	\frac{C_n^2}{n(C^{\star})^2} & = \frac{C_n^2n^{2s/d}}{c^2n(C_nn^{s/d})^{8s/(4s+d)}} \\
+	& = \frac{C_n^{2d/(4s + d)}n^{c^2(s/d)2d/(4s + d)}}{n} \\
+	& = \frac{(C_n n^{s/d})^{2d/(4s + d)}}{c^2n}
 	\end{align*}
-	and therefore $\epsilon \geq 2b\sqrt{2N(C^*)/n^2} + C_n^2/n(C^*)^2$. As a result, by Lemma \ref{lem:type_II} for any $\beta \in \mathcal{S}^s_d(C_n)$ such that $\norm{\beta}/n \geq \epsilon$, 
+	and therefore $\epsilon \geq 2b\sqrt{2N(C^*)/n^2} + C_n^2/n(C^*)^2$. As a result, by Lemma \ref{lem:type_II} for any $\beta \in \mathcal{S}^s_d(C_n)$ such that $\norm{\beta}/\sqrt{n} \geq \epsilon$, 
 	\begin{equation*}
-	\Ebb_{\beta}(1 - \phi_{C^{\star}}) \geq \frac{2}{b^2} + \frac{2}{b\sqrt{2N(C^{\star})}}.
+	\Ebb_{\beta}(1 - \phi_{C^{\star}}) \leq \frac{2}{b^2} + \frac{2}{b\sqrt{2N(C^{\star})}}.
 	\end{equation*}
@@
 
-In the following result, we prove that for an appropriate choice of threshold $C$ and radius $r$, the test $\phi_C$ achieves (nearly) minimax-optimal rates. Let
+In the following result, we prove that for an appropriate choice of threshold $C$ and radius $r$, the test $\phi_C$ achieves (nearly) minimax-optimal rates, whenever $s > d/4$.
+(I.e whenever $\mathcal{W}^{s,2}(\mathcal{D})$ can be embedded into $L^4(\mathcal{D})$.
+Let
+\begin{equation}
+\label{eqn:C_star_continuous}
+C^{\star} = \frac{\left(n^{1 + s/d}r^{d/2 + s}\right)^{4s/(4s + d)}}{n^{s/d}}.
+\end{equation}
+and suppose that $P$ be an absolutely continuous probability measure over $\D$ with density function $p$ bounded above and below by constants, i.e
 \begin{equation*}
-C^{\star} = \frac{\left(n^{1 + s/d}r^{d/2 + s}\right)^{4s/(4s + d)}}{n^{s/d}}.
+0 < p_{\textrm{min}} < p(x) < p_{\textrm{max}} < \infty, \quad \textrm{for all $x \in \mathcal{D}$.}
 \end{equation*}
@@
 	\label{thm:sobolev_testing_rate}
-	Let $b \geq 1$ be a fixed constant, and let $s = 1$. For any sequence $r_n$ such that $(\log n/n)^{1/d}r \to \infty$, there exists a constant $c$ which depends only on $d$ and $L$ such that
+	Let $b \geq 1$ be a fixed constant, let $s = 1$, and let $4s < d$.
+For any sequence $r_n$ such that $(\log n/n)^{1/d}r_n \to \infty$, and $(n^{1 + s/d}r_n^{d/2 + s})^{4s/(4s + d)}n^{-s/d} \leq \sqrt{2}$ for all $n$, there exists a constant $c_1$ which depends only on $d$, $L$, $p_{\max}$, and $b$ such that
 	\begin{equation}
 	\label{eqn:sobolev_testing_rate_1}
-	\mathcal{R}_{\epsilon}(\phi_{C^{\star}}; \mathcal{W}^{s,2}(\mathcal{D};L)) \leq c_1\left(\frac{1}{b^2} + \frac{1}{b\sqrt{N(C^{\star})}}\right) + o(1)
+	\mathcal{R}_{\epsilon}(\phi_{C^{\star}}; \mathcal{W}^{s,2}(\mathcal{D};L)) \leq \left(\frac{2}{b^2} + \frac{2}{b\sqrt{N(C^{\star})}}\right) + o(1)
 	\end{equation}
@@
 	\label{eqn:sobolev_testing_rate}
-	\epsilon^2 \geq c_2 b \left(\frac{(b n^{1 + s/d}r^{d/2 + s})^{2d/(4s + d)}}{n}\right).
+	\epsilon^2 \geq c_1 b \left(\frac{(b n^{1 + s/d}r^{d/2 + s})^{2d/(4s + d)}}{n}\right).
 	\end{equation}
 \end{theorem}
+\begin{remark}
+	Observe that when $4s > d$, the inequality $(n^{1 + s/d}r^{d/2 + s})^{4s/(4s + d)}n^{-s/d} \leq \sqrt{2}$ is satisfied whenever $r_n \leq n^{-3/(2d + 4s)} >> (\log(n)/n)^{1/d}$.
+In this regime, therefore, there exists a range of $r$ such that both conditions required in Theorem \ref{thm:sobolev_testing_rate} are satisfied.
+In particular, in this regime, choosing $r = \log(n)^{\alpha} (\log(n)/n)^{1/d}$ for any $\alpha > 0$, we have that the lower bound in \eqref{eqn:sobolev_testing_rate} becomes
+	\begin{equation*}
+	c_1 b (\log n)^{(1/d + \alpha)(d/2 + s)(2d/(4s + d))} n^{-4s/(4s + d)}
+	\end{equation*}
+	matching, up to log factors, the minimax rate for testing over Sobolev functions.
+\end{remark}
+\begin{remark}
+	Many rates of nonparametric hypothesis testing exhibit an elbow when $4s < d$.
+In our particular setting, it can be shown that a naive test based on the statistic $\norm{z}_2^2$ has power for alternatives with norm $\norm{f}^2 \geq n^{-1/4}$, regardless of $s$ and $d$, for any $f \in L^4(\mathcal{D})$.
+\end{remark}
 
@@
 \begin{equation*}
-\mathcal{R}_{\epsilon}(\phi_{C^{\star}}; W^{1,2}(\mathcal{D};L)) = \Ebb_{f = 0}(\phi_{C^{\star}}) + \sup_{f \in W^{1,2}(\mathcal{D};L): \norm{f}_2 > \epsilon} \Ebb_{f}(1 - \phi_{C^{\star}})
+\mathcal{R}_{\epsilon}(\phi_{C^{\star}}; W^{1,2}(\mathcal{D};L)) = \Ebb_{f = 0}(\phi_{C^{\star}}) + \sup_{f \in W^{1,2}(\mathcal{D};L): \norm{f}_{P,2} > \epsilon} \Ebb_{f}(1 - \phi_{C^{\star}})
 \end{equation*}
@@
 	\label{event:discrete_sobolev_norm}
-	Discrete Sobolev norm of $\beta$: There exists a constant $c_1$ which depends only on $p_{\max}$ and $d$ such that
+	Discrete Sobolev norm of $\beta$: There exists a constant $c_2$ which depends only on $p_{\max}$ and $d$ such that
 	\begin{equation*}
-	\norm{B^{(s)}\beta}_2^2 \leq c_1 b^2 n^2 r^{d + 2s} 
+	\norm{B^{(s)}\beta}_2^2 \leq c_2 b^2 n^2 r^{d + 2s} 
 	\end{equation*}
@@
 Each of these events happen with high probability as $n \to \infty$.
-\begin{lemma}
+\begin{lemma}[High Probability Events.]
 	\label{lem:high_probability_events}
-	With probability at least $1 - \frac{1}{b^2}$ as $n \to \infty$,
+	\mbox{}
+
+	\textbf{Discrete Sobolev norm.}
+	There exists a constant $c_2$ which depends only on $p_{\max}$ and $d$ such that, with probability at least $1 - \frac{1}{b^2}$,
 	\begin{equation}
 	\label{eqn:discrete_sobolev_norm}
-	\norm{B\beta}_2^2 \leq c_1 b^2 n^2 r^{d + 2}.
+	\norm{B\beta}_2^2 \leq c_2 b^2 n^2 r^{d + 2}.
 	\end{equation}
 
-	With probability tending to one as $n \to \infty$,
+	\textbf{Eigenvalue tail decay.}
+	Suppose the sequence of neighborhood graph radii $(r_n)$ satisfies
+	\begin{itemize}
+		\item $r(n/\log n)^{1/d} \to \infty$, and
+		\item $r_n > \sqrt{2}n^{-1/(2d + 4s)}$ for all $n$.
+	\end{itemize} 
+	Then with probability tending to one as $n \to \infty$,
 	\begin{equation}
@@
 
-	\textcolor{red}{$L_2$ norm}.
+	\textbf{Norm of $\beta$.}
+	There exists a constant $c_5$ which depends only on $s,d$, and $\mathcal{D}$ such that
+	\begin{equation}
+	\label{eqn:l2_norm}
+	\frac{\norm{\beta}^2}{n} < \frac{\epsilon^2}{2}
+	\end{equation}
+	with probability at least $1  - 4c_5p_{\max}^4L^4/(n\epsilon^2) \to 0$ as $n \to \infty$. 
 \end{lemma}
-In the following subsections, we prove Lemma \ref{lem:high_probability_events}.
-First, we show that Theorem \ref{thm:sobolev_testing_rate} is a simple consequence of Lemma \ref{lem:high_probability_events} along with our previous work.
+
+In the following subsections, we prove the various statements which make up Lemma \ref{lem:high_probability_events}.
+First, however, we show that Theorem \ref{thm:sobolev_testing_rate} is a simple consequence of Lemma \ref{lem:high_probability_events} along with the fixed graph testing results of Section \ref{sec:fixed_graph_testing}.
 Let $\mathcal{A}_f$ be the set of $X$ such that \ref{event:discrete_sobolev_norm}, \ref{event:eigenvalue_tail_decay}, and \ref{event:l2_norm} all occur, for the given choice of $f$.
-We have that for $C_n := c_1bnr^{d/2 + s}$, the event \ref{event:discrete_sobolev_norm} implies $\beta \in \mathcal{S}_d^s(C_n)$.
+We have that for $C_n := \sqrt{c_2}bnr^{d/2 + s}$, the event \ref{event:discrete_sobolev_norm} implies $\beta \in \mathcal{S}_d^s(C_n)$.
 Therefore, by \ref{event:l2_norm},
@@
 \begin{itemize}
-	\item \textcolor{red}{$C^{\star} = (C_n n^{s/d})^{4s/(4s+d)}/n^{s/d}$},
-	\item The event \ref{event:eigenvalue_tail_decay} is exactly \ref{asmp:tail_decay}, and 
-	\item For an appropriate choice of $c_2$, by hypothesis
+	\item $C^{\star} = c_3(C_n n^{s/d})^{4s/(4s+d)}/n^{s/d}$, where $c_3 = (b c_2^{1/2})^{-1}$. 
+	\item The event \ref{event:eigenvalue_tail_decay} is exactly \ref{asmp:tail_decay}, specialized to our choice of graph \textcolor{red}{$G_{n,r}$}, and 
+	\item Letting $c_1 := 2c_2^{-2d/(4s + d)}\left(2bc_3^{d/2s} + c_3^{-2}\right)$, by \eqref{eqn:sobolev_testing_rate},
 	\begin{equation*}
-	\epsilon^2/2 \geq (2\sqrt{2}b + 1) \left(\frac{(C_n n^{s/d})^{2d/(4s + d)}}{n}\right)
+	\epsilon^2/2 \geq (2bc_3^{d/2s} + c_3^{-2}) \left(\frac{(C_n n^{s/d})^{2d/(4s + d)}}{n}\right)
 	\end{equation*}
@@
 
+\begin{lemma}
+	\label{lem:sobolev_1_bound}
+	For any $f \in W^{1,2}(\mathcal{D};L)$, and any $b \geq 1$, we have that there exists a constant $c_2 > 0$ which depends only on $d$ and $p_{\max}$ such that
+	\begin{equation}
+	\label{eqn:sobolev_1_bound}
+	\norm{B\beta}_2^2 \leq L^2 b^2 c_2 n^2 r^{d + 2}
+	\end{equation}
+	with probability at least $1 - \frac{1}{b^2}$. 
+\end{lemma}
+\begin{proof}
+	Observe that
+	\begin{align*}
+	\frac{1}{n^2}\Ebb(\beta^T L \beta) = \Ebb\left(\frac{1}{n^2} \sum_{i,j = 1}^{n} (\beta_i - \beta_j)^2 A_{ij}\right) & = \frac{(n - 1)}{n} \int_{\D} \int_{\D} (f(x) - f(y))^2\eta_r(x,y) \,dP(y) \,dP(x) \\
+	& \leq p_{\max}^2 \frac{(n - 1)}{n} \int_{\D} \int_{\D} (f(x) - f(y))^2\eta_r(x,y) \,dy \,dx.
+	\end{align*}
+	We will show that for any $f \in \mathcal{W}^{1,2}(\D;L)$, there exists a constant $c_4$ which depends only on dimension $d$ such that
+	\begin{equation}
+	\label{eqn:mean_bound}
+	\int_{\D} \int_{\D} (f(x) - f(y))^2\eta_r(x,y) \,dy \,dx \leq c_4 L^2 r^{d + 2}
+	\end{equation}
+	whence the desired result of \eqref{eqn:sobolev_1_bound} follows by Markov's inequality.
+
+	We begin by dealing with complications due to the boundary of $\mathcal{D}$.
+Let $V$ be any bounded open set such that $\D \subset \subset V$.
+Note that as $\partial \D$ is $C^1$, by Theorem \ref{thm:evans_extension} there exists $g \in W^{1,2}(\Reals^d)$ such that
+	\begin{enumerate}
+		\item
+		\label{eqn:evans_extension_1}
+		$g = f$, $P$-almost-everywhere in $\D$
+		\item 
+		$g$ has support within $V$, and  
+		\item 
+		\label{eqn:sobolev_1_bound_2}
+		$\norm{g}_{W^{1,2}(\Rd)} \leq C \norm{f}_{W^{1,2}(\D)}$ for a constant $c$ which depends only on $\mathcal{D}$.
+	\end{enumerate}
+	As a result of the first point,
+	\begin{equation}
+	\label{eqn:sobolev_1_bound_1}
+	\int_{\D} \int_{\D} (f(x) - f(y))^2\eta_r(x,y) \,dy \,dx \leq \int_{\Rd} \int_{\Rd} (g(x) - g(y))^2\eta_r(x,y) \,dy \,dx.
+	\end{equation}
+
+	Next, we smooth $g$, so that we may work with ordinary partial derivatives.
+	We let $\kappa \in C^{\infty}(\Rd)$ be given by
+	\begin{equation*}
+	\kappa(x) :=
+	\begin{cases}
+	C \exp \left\{\frac{1}{\norm{x}^2 - 1}\right\}& \quad \textrm{if $\norm{x}_2 \leq 1$} \\
+	0 & \quad \textrm{if $\norm{x}_2 \geq 1$}
+	\end{cases}
+	\end{equation*}
+	where the normalizing constant $C > 0$ is chosen so that $\int_{\Rd} \eta dx = 1$. Let $\kappa_r(x) := (1/r^d) \kappa(x/r)$. Then, the mollification of $g$ by $\kappa_r$ is given by
+	\begin{align*}
+	g^r & := g \ast \eta_r \\
+	& = \int_{\Rd} \eta_r(x - y) g(y) dy
+	\end{align*}
+	(Refer to \citep{evans10}, Appendix C, Theorem 7 for a proof that $g^r \in C^{\infty}(\Rd)$.)
+	Adding and subtracting within \eqref{eqn:sobolev_1_bound_1}, we have
+	\begin{align}
+	\int_{\Rd} & \int_{\Rd} (g(x) - g(y))^2\eta_r(x,y) \,dy \,dx \\
+	& \leq 3 \int_{\Rd} \int_{\Rd} \bigl((g(x) - g^r(x))^2 + (g^r(x) - g^r(y))^2 + (g^r(y) - g(y))^2\bigr)\eta_r(x,y) \,dy \,dx \nonumber \\
+	& = 6 \int_{\Rd} \int_{\Rd} \bigl((g(y) - g^r(y))^2 \eta_r(x,y) \,dy \,dx + 2 \int_{\Rd} \int_{\Rd} \bigl((g^r(x) - g^r(y))^2 \eta_r(x,y) \,dy \,dx \label{eqn:sobolev_1_bound_5}
+	\end{align}
+	We deal with each summand individually, beginning with the first one. Let $\nu_d = \pi^{d/2}/\Gamma(d/2 + 1)$ denote the volume of a ball with unit radius in $\Rd$. Then, we have
+	\begin{align}
+	\int_{\Rd} \int_{\Rd} \bigl((g(y) - g^r(y))^2 \eta_r(x,y) \,dy \,dx & = \int_{\Rd} \int_{B(x,r)} \bigl((g(y) - g^r(y))^2\,dy \,dx \nonumber \\
+	& = \int_{\Rd} \int_{B(y,r)} \bigl((g(y) - g^r(y))^2\,dx \,dy \tag{Tonelli's Theorem} \nonumber \\
+	& = \nu_d r^d \int_{\Rd}\bigl((g(y) - g^r(y))^2 \,dy \\
+	& \leq \nu_d r^{d + 2} \int_{\Rd} \norm{\nabla g(y)}^2 \,dy \label{eqn:sobolev_1_bound_4}
+	\end{align}
+	where the last line follows from Lemma \ref{lem:poincare_mollify}.
+
+	We now turn our attention to the second summand. Note that as $g^r \in C^{\infty}(\Rd)$, we may apply Theorem \ref{thm:taylor_expansion} and obtain
+	\begin{align*}
+	(g^r(x) - g^r(y))^2 & = \left(\int_{0}^{1} \nabla g^r(x + t(y - x)) \cdot (y - x) \,dt\right)^2 \\
+	& \leq \int_{0}^{1} \bigl(\nabla g^r(x + t(y - x)) \cdot (y - x)\bigr)^2 \,dt  \tag{Jensen's inequality} \\
+	& \leq \norm{y - x}^2 \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 \,dt. \tag{Cauchy-Schwarz inequality}
+	\end{align*}
+	As a result, we have
+	\begin{align*}
+	\int_{\Rd} \int_{\Rd} \bigl((g^r(x) - g^r(y))^2 \eta_r(x,y) \,dy) \,dx & \leq \int_{\Rd} \int_{\Rd} \norm{y - x}^2  \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 \eta_r(x,y) \,dt \,dy \,dx \\
+	& \leq r^{2} \int_{\Rd} \int_{\Rd} \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 \eta_r(x,y) \,dt \,dy \,dx \\
+	& = r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + t(y - x))}^2 \eta_r(x,y) \,dx \,dt \,dy \\
+	& = r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + tz)}^2 \eta_r(z) \,dx \,dt \,dz \tag{$z = y - x$}
+	\end{align*}
+
+	where we write $\eta_r(z) = \mathbf{1}(\norm{z} \leq r)$ in an abuse of notation. Next, we note that
+	\begin{equation*}
+	\int_{\Rd} \norm{\nabla g^r(x + tz)}^2 \,dx = \int_{\Rd} \norm{\nabla g^r(x)}^2 \,dx \leq \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
+	\end{equation*}
+	with the inequality following from Lemma \ref{lem:gradient_mollify_commute}. Therefore,
+	\begin{align}
+	r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + tz)}^2 \eta_r(z) \,dx \,dt \,dz & \leq r^{2} \int_{\Rd} \eta_r(z) \int_{0}^{1} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx  \,dt \,dz \nonumber \\
+	& = \nu_d r^{2 + d} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx. \label{eqn:sobolev_1_bound_3}
+	\end{align}
+	By \eqref{eqn:sobolev_1_bound_5}, \eqref{eqn:sobolev_1_bound_4} and \eqref{eqn:sobolev_1_bound_3}, we have that 
+	\begin{equation*}
+	\int_{\Rd} \int_{\Rd} (g(x) - g(y))^2 \eta_r(x,y) \,dx \,dy \leq 9 \nu_d r^{d + 2} \int_{\Rd} \norm{\nabla g(x)}_2^2 \,dx
+	\end{equation*}
+	Then by \eqref{eqn:sobolev_1_bound_2}, $\int_{\Rd} \norm{\nabla g(x)}_2^2 \,dx \leq c \int_{\D} \norm{\nabla f(x)}_2^2 \,dx \leq c L^2 $ where $c$ is a constant depending only on $\D$.
+So the desired result of \eqref{eqn:sobolev_1_bound_1} follows.
+\end{proof}
+
+\subsection{Proof of \eqref{eqn:eigenvalue_tail_decay}.}
+
+\begin{theorem}
+	\label{thm:spectral_decay}
+	Let $r \to 0$ as $n \to \infty$ sufficiently slowly so that $r(\log n/n)^{1/d} \to \infty$. Then, for each $s > 0$ and $C \leq \sqrt{2}$,
+	\begin{equation}
+	\label{eqn:spectral_decay}
+	N(C) \leq n C^{d/s}.
+	\end{equation} 
+	with probability tending to one as $n \to \infty$. 
+\end{theorem}
+
+Before we prove Theorem \ref{thm:spectral_decay}, we observe that, whenever $4s > d$, the choice of $C^{\star}$ by \eqref{eqn:C_star_continuous} along with the requirement that $r_n < \sqrt{2}(n^{-3/(2d + 4s)})$ implies $C^{\star} < \sqrt{2}$ for all $n$.
+Therefore, Theorem \ref{thm:spectral_decay} implies \eqref{eqn:eigenvalue_tail_decay}, the desired result.
+
+To prove Theorem \ref{thm:spectral_decay}, we will make use of a grid graph $\widetilde{G}$ on points within $\mathcal{D}$, whose spectral properties are very well-understood.
+Let $\xi$ be the set of evenly spaced lattice points over $\mathcal{D}$; formally $\xi = \set{k/n: k \in [\ell]^d}$ where $\ell = n^{1/d}$, and we define $[\ell]^d = \set{(\ell_1,\ldots,\ell_d): \ell_k \in [\ell] ~\textrm{for each}~ \ell_k}$.
+Then, let the grid graph over $\xi$ be given by $\wt{G} = (\wt{V},\wt{E})$, where $\wt{V} = \xi$ and $(\xi_k, \xi_{k'})$ is in $\wt{E}$ if $\norm{\xi_k - \xi_{k'}}_1 = 1/\ell$.
+
+Let $\wt{L}$ be the Laplacian of $\wt{G}$, and let $\wt{\lambda_1} \leq \ldots \leq \wt{\lambda}_n$ be the ordered eigenvalues of $\wt{L}$.
+Write $\wt{N}(C) = \#\set{k: \wt{\lambda}_k^s \leq C^2}$.
+We have that the desired scaling rate \eqref{eqn:spectral_decay} holds with respect to the grid graph $\wt{G}$.
+
+\begin{lemma}
+	\label{lem:spectral_decay_grid}
+	For each $s$ and every $C \leq \sqrt{2}$, we have
+	\begin{equation}
+	\label{eqn:spectral_decay_grid}
+	\wt{N}(C) \leq 2^d\big(n C^{d/s} + 1)
+	\end{equation}
+\end{lemma}
+\begin{proof}
+	It will be sufficient to show that 
+	\begin{equation}
+	\label{eqn:spectral_decay_grid_1}
+	\lambda_k^s \leq C^2 \Rightarrow \floor{k^{1/d}}^d \leq C^{d/s}n
+	\end{equation}
+	To show this, observe that for any $\tau \in \mathbb{N}$ and $k = \tau^d$, we have
+	\begin{equation*}
+	\lambda_k \geq 4 \sin^2 \left(\frac{\pi k^{1/d}}{2n^{1/d}}\right) \geq \frac{\pi^2 k^{2/d}}{4 n^{2/d}} \wedge 2
+	\end{equation*}
+	Therefore, if $\lambda_k^s \leq C^2$ and $C \leq \sqrt{2}$, this implies
+	\begin{equation*}
+	\frac{\pi^{2s} k^{2s/d}}{4^s n^{2s/d}} \leq C^{2}
+	\end{equation*}
+	and rearranging, we obtain
+	\begin{equation*}
+	k \leq \frac{C^{d/s} 2^d n}{\pi^d} \leq C^{d/s}n
+	\end{equation*}	
+	If $k^{1/d}$ is not a natural number, applying the same argument to $k' = \floor{k^{1/d}}^d$ yields \eqref{eqn:spectral_decay_grid_1}.
+\end{proof}
+
+In light of Lemma \ref{lem:spectral_decay_grid}, to prove Theorem \ref{thm:spectral_decay} it is sufficient to show that $\wt{L} \preceq L$, since by the Courant-Fischer min-max theorem, the ordering $\wt{L} \preceq L$ implies that $\wt{\lambda}_k \leq \lambda_k$ for all $k \in [n]$.
+The next result details the conditions under which this ordering holds.
+This condition will be stated with respect to the min-max matching distance between $\xi$ and $X$, i.e.
+the minimum over all bijections $T: \xi \to X$ such that
+\begin{equation*}
+\min_{T} \max_{i \in [n]} \abs{T^{-1}(x_i) - x_i} 
+\end{equation*}
+
+\begin{lemma}
+	\label{lem:partial_ordering_grid}
+	For any radius $r$ satisfying
+	\begin{equation}
+	\label{eqn:radius_condition}
+	r \geq 2 \min_{T} \max_{i \in [n]} \abs{T^{-1}(x_i) - x_i} + n^{-1/d}
+	\end{equation}
+	we have that $\wt{L} \preceq L$. 
+\end{lemma}
+\begin{proof}
+	Let $T_{\star}$ achieve the min-max matching distance, i.e
+	\begin{equation*}
+	\max_{i \in [n]} \abs{T_{\star}^{-1}(x_i) - x_i} = \min_{T} \max_{i \in [n]} \abs{T^{-1}(x_i) - x_i}.
+	\end{equation*}
+
+	It will be sufficient to prove that for every pair $(T_{\star}^{-1}(x_i), T_{\star}^{-1}(x_j)) \in \wt{E}$, the corresponding edge $(i,j)$ is in $E$.
+To see this, let $A$ denote the adjacency matrix associated with the neighborhood graph $G$, and $\wt{A}$ the adjacency matrix associated with the grid $\wt{G}$.
+Precisely $A$ is the $n \times n$ matrix with entries $A_{ij} = \eta_r(x_i,x_j)$, and $\wt{A}$ is the $n \times n$ matrix with entries $\wt{A}_{ij} = \1\{T_{\star}^{-1}(x_i), T_{\star}^{-1}(x_j) \in \wt{E} \}$.
+Our goal is to show that, for every $z \in \Rd$, we have
+	\begin{equation*}
+	z^T L z = \frac{1}{2}\sum_{i, j = 1}^{n} (z_i - z_j)^2 A_{ij} \leq \frac{1}{2}\sum_{i, j = 1}^{n} (z_i - z_j)^2 \wt{A}_{ij} = z^T \wt{L} z.
+	\end{equation*}
+	which certainly holds if $\wt{A}_{ij} \leq A_{ij}$ for all $i,j \in [n]$.
+
+	Now, assume $(T_{\star}^{-1}(x_i), T_{\star}^{-1}(x_j)) \in \wt{E}$. This implies
+	\begin{align*}
+	\norm{x_i - x_j}_2 & \leq \norm{x_i - T_{\star}^{-1}(x_i)}_2 + \norm{T_{\star}^{-1}(x_j) - T_{\star}^{-1}(x_j)}_2 + \norm{x_j - T_{\star}^{-1}(x_j)}_2 \\
+	& \leq 2 \max_{i \in [n]} \abs{T_{\star}^{-1}(x_i) - x_i} + n^{-1/d} \\
+	& \leq r.
+	\end{align*}
+	so $\eta_r(x_i,x_j) = 1$ and therefore $(i,j) \in E$. 
+\end{proof}
+
+Finally, the following Lemma demonstrates that, for sufficiently large $r$, the condition \eqref{eqn:radius_condition} will hold with high probability.
+
+\begin{lemma}
+	\label{lem:matching_distance}
+	Assume $P$ has density $p$ which is bounded above and below uniformly over $\D$; that is, there exist constants $p_{\min}$ and $p_{\max}$ such that
+	\begin{equation*}
+	0 < p_{\min} < p(x) < p_{\max} < \infty, \quad \text{for all $x \in \mathcal{D}$}.
+	\end{equation*}
+	Then, for any $r = r_n$ such that $r(\log(n)/n)^{1/d} \to \infty$, we have that the event
+	\begin{equation*}
+	r < 2 \min_{T} \max_{i \in [n]} \abs{T^{-1}(x_i) - x_i} + n^{-1/d}
+	\end{equation*}
+	occurs with probability tending to $0$ as $n \to \infty$. 
+\end{lemma}
+
+Together Lemmas \ref{lem:spectral_decay_grid}, \ref{lem:partial_ordering_grid}, and \ref{lem:matching_distance} imply Theorem \ref{thm:spectral_decay}.
+
+\subsection{Proof of \eqref{eqn:l2_norm}}
+
+Rewriting $\norm{\beta}^2/n = n^{-1} \sum f(x_i)^2$, we use Chebyshev's inequality to obtain
+\begin{align*}
+\Pbb\Bigl(n^{-1} \sum_{i = 1}^{n} f(x_i)^2 < \epsilon^2/2\Bigr) & \leq \Pbb\left( \left(n^{-1} \sum_{i=1}^{n} f(x_i)^2 - \int_{\D} f^2 \,dP(x)\right)^2 > \frac{\left(\int_{\D} f^2 \,dP(x)\right)^2}{4}\right) \\
+& \leq 4 \frac{\Var\left(f^2(x_1)\right)}{n(\int f^2 dP)^2} \\
+& \leq 4 \frac{\int f^4 \,dP(x) }{n\epsilon^2} \\
+& \leq 4 p_{\max}^4 \frac{\int f^4 \,dx }{n\epsilon^2}
+\end{align*}
+Then, since we have assumed $4s > d$, by the Sobolev embedding theorem we have that there exists some $c_5$ depending only on $s,d$, and $\D$ such that
+\begin{equation*}
+\int f^4 \,dx = \norm{f}^4_{L^4(\D)} \leq c \norm{f}_{W^{2,s}(\mathcal{D})}^4.
+\end{equation*}
+\section{Supporting Results.}
+
+\begin{theorem}[\citep{evans10} Chapter 5.4, Theorem 1]
+	\label{thm:evans_extension}
+	Assume $U$ is bounded and $\partial U$ is $C^1$.
+Select a bounded open set $V$ such that $U \subset \subset V$ ($U$ is compactly contained in $V$).
+Then there exists a bounded linear operator $E: W^{1,2}(U) \to W^{1,2}(\Rd)$ such that for each $u \in W^{1,2}(U)$:
+	\begin{enumerate}
+		\item $Eu = u$ a.e. in $U$,
+		\item $Eu$ has support within $V$, and 
+		\item 
+		\begin{equation*}
+		\norm{Eu}_{W^{1,2}(\Rd)} \leq C \norm{u}_{W^{1,2}(\Rd)}
+		\end{equation*}
+		the constant $C$ depending only on $U$ and $V$.
+	\end{enumerate}
+\end{theorem}
+
+For $u \in W^{1,2}(\Rd)$ and $x \in \Rd$, write $\nabla u(x) = (D^{e_1}(x),\ldots,D^{e_d}(x)$, for the gradient of $u$.
+
+\begin{theorem}[\citep{evans10} Chapter 5.8.1, Theorem 2]
+	\label{thm:evans_poincare}
+	There exists a constant $C$, depending only on $d$, such that
+	\begin{equation*}
+	\norm{u - u^r}_{L^2(B(x,r))} \leq Cr\norm{\nabla u}_{L^2(B(x,r))}
+	\end{equation*}
+\end{theorem}
+
+\begin{theorem}[Taylor expansion.]
+	\label{thm:taylor_expansion}
+	For any function $u \in C^1$, and any $x,y \in \Rd$, 
+	\begin{equation*}
+	u(y) - u(x) = \int_{0}^{1} \nabla(u(x + t(y - x))) \cdot (y - x) \,dt
+	\end{equation*}
+\end{theorem}
+
+\begin{lemma}
+	\label{lem:poincare_mollify}
+	For any function $g \in W^{1,2}(\Rd)$ compactly supported in a bounded open set $V \subset \Rd$, we have
+	\begin{equation}
+	\label{eqn:poincare_mollify}
+	\int_{\Rd} (g(x) - g^r(x))^2 \,dx \leq r^2 \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
+	\end{equation}
+\end{lemma}
+\begin{proof}
+	This Lemma is essentially a reproduction of part of the proof of the Rellich-Kondrachov Compactness Theorem from \citep{evans10}.
+Note that it is sufficient to prove in the case when $g$ is smooth.
+To see this, for the moment assume \eqref{eqn:poincare_mollify} holds for all $u \in C^{\infty}(V)$, and let $g \in W^{1,2}(V)$.
+By Theorem \ref{thm:local_approx_smooth_functions}, we may take a sequence $g_m \in C^{\infty}(V)$ such that
+	\begin{equation*}
+	\norm{g - g_m}_{L^2(V)} \to 0, \quad \textrm{and} \quad \norm{\nabla g - \nabla g_m}_{L^2(V)} \to 0,
+	\end{equation*}
+	Then we have
+	\begin{align*}
+	\int_{\Rd} (g(x) - g^r(x))^2 \,dx & \leq \int_{\Rd} (g(x) - g_m(x))^2 \,dx + \int_{\Rd}(g_m(x) - g_m^r(x))^2 \,dx + \int_{\Rd}(g_m^r(x) - g^r(x))^2 \,dx \\
+	& \leq \int_{\Rd} (g(x) - g_m(x))^2 \,dx + r^2 \int_{\Rd} \norm{\nabla g_m(x)}^2 \,dx + \int_{\Rd}(g_m^r(x) - g^r(x))^2 \,dx
+	\end{align*}
+	and taking the limit as $m$ goes to infinity, the right hand side converges to $\int_{\Rd} \norm{\nabla g(x)}^2 \,dx$.
+
+	It remains to show \eqref{eqn:poincare_mollify} in the case where $g$ is smooth. In this case, we have
+	\begin{align*}
+	g^{r}(x) -  g(x) & = \frac{1}{r^d} \int_{B(x,r)} \kappa \left(\frac{x - z}{r}\right)\bigl(g(z) - g(x)\bigr) \,dz \\
+	& = \int_{B(0,1)} \kappa(y) \bigl(g(x - ry) - g(x)\bigr) \,dy \\
+	& = \int_{B(0,1)} \kappa(y) \int_{0}^{1} \frac{d}{dt} \bigl(g(x - try) \bigr) \,dt \,dy \\
+	& = -r \int_{B(0,1)} \kappa(y) \int_{0}^{1} \bigl(\nabla g (x - try) \bigr) \cdot y \,dt \,dy.
+	\end{align*} 
+	Therefore, by Jensen's and Cauchy-Schwarz inequalities, we have
+	\begin{align*}
+	\int_{\Rd} (g(x) - g^r(x))^2 \,dx & \leq r^2 \int_{\Rd} \int_{B(0,1)} \kappa(y) \int_{0}^{1} \norm{\nabla g (x - try)}^2 \norm{y}^2 \,dt \,dy \,dx \\
+	& \leq r^2 \int_{B(0,1)} \kappa(y) \int_{0}^{1} \int_{\Rd} \norm{\nabla g (x - try)}^2 \,dx \,dt \,dy \\
+	& = r^2 \int_{\Rd} \norm{\nabla g(z)}^2 \,dz
+	\end{align*}
+\end{proof}
+
+The following theorem is Theorem 1 in Section 5.3 of \citep{evans10}.
+\begin{theorem}[Local approximation by smooth functions.]
+	\label{thm:local_approx_smooth_functions}
+	Assume $U$ is bounded, and $u \in W^{k,p}(U)$ for some $1 \leq p < \infty$. Then there exists functions $u_m \in C^{\infty}(U) \cap W^{k,p}(U)$ such that
+	\begin{equation*}
+	\norm{u_m - u}_{W^{k,p}(U)} \overset{m}{\to} 0.
+	\end{equation*}
+\end{theorem}
+
+\begin{lemma}
+	\label{lem:gradient_mollify_commute}
+	For any $u \in W^{1,2}(\Rd)$, we have
+	\begin{equation*}
+	\int_{\Reals^d} \norm{\nabla u^r(x)}_2^2 \,dx \leq \int_{\Reals^d} \norm{\nabla u(x)}_2^2 \,dx
+	\end{equation*}
+\end{lemma}
+\begin{proof}
+	Observe that $\norm{\nabla u(x)}_2^2 = \sum_{j = 1}^{d} (D^{e_j}u(x))^2 $. Therefore,
+	\begin{align*}
+	\int_{\Rd} \norm{\nabla u^r(x)}_2^2 \,dP(x) & = \sum_{j = 1}^{d} \int_{\Rd} (D^{e_j}u^r(x))^2 \,dx \\
+	& = \sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u)^r(x))^2 \,dx
+	\end{align*}
+	where the second equality follows from equation (1) in Section 5.3 of \citep{evans10}. Then, for any $v \in L^2(\Rd)$, we have that
+	\begin{align*}
+	\abs{v^r(x)} & = \int_{\Rd} \kappa_r^{1/2}(x - y) \kappa_r^{1/2}(x - y) v(y) \,dy \\
+	& \leq \left(\int_{\Rd} \kappa_r^(x - y) \,dy\right)^{1/2} \left(\int_{\Rd} \kappa_r(x - y) v^2(y) \right)^{1/2} \\
+	& = \left(\int_{\Rd} \kappa_r(x - y) v^2(y) \right)^{1/2}
+	\end{align*}
+	and therefore
+	\begin{align*}
+	\int_{\Rd} (v^r(x))^2 \,dx & \leq \int_{\Rd} \int_{\Rd } \kappa_r(x - y) v^2(y) \,dy \,dx \\
+	& = \int_{\Rd} \int_{\Rd } \kappa_r(x - y) v^2(y) \,dy \,dx \\
+	& = \int_{\Rd} v^2(y) \,dy
+	\end{align*}
+
+	Applying this to $D^{e_j}u \in L^2(\Rd)$, we have that
+	\begin{equation*}
+	\sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u)^r(x))^2 \,dx \leq \sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u(x))^2 \,dx = \int_{\Rd} \norm{\nabla u(x)}^2 \,dx.
+	\end{equation*}
+\end{proof}
+
+\bibliographystyle{plainnat}
+\bibliography{../graph_testing_bibliography}
+
 \end{document}

### 2019-11-18 notes/proposal/proposal_document.tex
added proposal draft

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\mathrm{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}

%%% Sets and classes
\newcommand{\Cset}{\mathcal{C}}
\newcommand{\Csig}{\Cset_{\sigma}}
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\domain}{\mathcal{X}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Cbb}{\mathbb{C}}

%%% Functionals
\newcommand{\cut}{\textrm{cut}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\Partial}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\dist}{\mathrm{dist}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Thesis Proposal}
\author{Alden Green}
\date{\today}
\maketitle

\begin{abstract}
	In the classical data analysis setting, where we observe i.i.d. point-cloud data sampled from an unknown distribution $P$, there has been recent interest in the design and analysis of procedures which involve the geometry of $P$. Broadly speaking, these procedures either estimate geometric properties such as high-density regions of $P$, or leverage the geometry of $P$ in some downstream statistical learning task. One way to implicitly adapt to the geometry of $P$ is through the use of graph-based spectral methods. The underlying intuition is that when one appropriately constructs a neighborhood graph, the resulting graph Laplacian encodes a great deal of information about the geometry of $P$; for example, one can view its top eigenvectors as (discrete versions of) functions which are smooth with respect to $P$. For this reason, there has been extensive analysis on consistency properties of graph Laplacians, that is, in what ways and with what error they approximate continuum-level operators.

	\vspace{.02 in}

	In this thesis we will take a different approach, and analyze the effectiveness with which spectral algorithms perform traditional statistical learning tasks such as density clustering and hypothesis testing. In the density clustering problem, we consider a local version of spectral clustering using the Personalized PageRank (PPR) algorithm. We show that when a density cluster satisfies a natural set of geometric conditions PPR will provably recover it with high probability; we also derive lower bounds conversely showing that PPR will fail to recover geometrically poorly conditioned density clusters. In the hypothesis testing context, in some preliminary work we define a goodness-of-fit test using spectral-based test statistics formed over a neighborhood graph, and prove that it achieves minimax optimal testing rates against a certain class of smooth alternatives. We propose various extensions of this test targeted for wider classes of alternatives, and for the two-sample problem.
\end{abstract}

\section{Background.}

\textcolor{red}{TODO:} This is why you are going to read.

\textcolor{red}{TODO:} This is what you are going to read.

\subsection{Neighborhood graphs and their spectra.}
\label{subsec:neighborhood_graphs_spectra}

Let $X = \{x_1,
\ldots, x_n\}$ be a sample drawn i.i.d.\ from a distribution $\Pbb$ on $\Rd$,
with density~$f$.  For a radius $r > 0$, we define $G_{n,r}=(V,E)$ to be the
\emph{$r$-neighborhood graph} of $X$, an unweighted, undirected graph with
vertices $V=X$, and an edge $(x_i,x_j) \in E$ if and only if $\norm{x_i -
	x_j} \leq r$, where $\norm{\cdot}$ is the Euclidean norm. We denote by $A \in
\Reals^{n \times n}$ the adjacency matrix, with entries $A_{uv} = 1$ if
$(u,v) \in E$ and $0$ otherwise.  We also denote by $D$ the diagonal degree
matrix, with entries $D_{uu} := \sum_{v \in V} A_{uv}$, and by $I$ the $n
\times n$ identity matrix.

Spectral graph theory involves the study of the eigenvalues and eigenvectors of any of following Laplacian matrices:
\begin{equation*}
\textrm{combinatorial:}~ L = D - A,~~ \textrm{random walk:}~ L_{\textrm{rw}} = I - D^{-1}A,~~ \textrm{normalized:}~ L_{\textrm{n}} = D^{1/2}L_{\textrm{rw}}D^{-1/2}.
\end{equation*}

and is broadly applicable to fields as diverse as the theory of randomized algorithms and differential geometry \textcolor{red}{Chung}. When specifically considering Laplacians of neighborhood graphs such as $G_{n,r}$, one well-developed line of work concerns the asymptotic convergence properties of the various graph Laplacians to associated continuum Laplacian operators.

\begin{itemize}
	\item  \textcolor{red}{Define the Laplacian operator}.
	\item Introduce the study of Laplacians on point-cloud data, manifold learning, and give references as to their background. Discuss the study of the convergence of spectra \textcolor{red}{Belkin, Coifman, Garcia-Trillos and Slepcev 1, Garcia-Trillos and Slepcev Spec, von Luxburg, Hein + von Luxburg, Lafon,Koltchinskii and Gine, Gine and Koltchinskii 1}, pointwise convergence of operators \textcolor{red}{Belkin thesis, Lafon thesis, Hein + Audibert, Singer, Belkin + Niyogi08}, and convergence of smoothness functionals \textcolor{red}{Bousquet + Chapelle, Hein07}. 
	\item Discuss the study of other functionals, such as total variation and normalized cut, on graphs \textcolor{red}{(Garcia-Trillos and Slepcev TV, Arias-Castro, Maier + von Luxburg)}
	\item 
\end{itemize}

Why should I care about the asymptotic convergence of graph Laplacians?
\begin{itemize}
	\item The convergence of graph Laplacians, and their spectra, to underlying continuum operators suggests they may be used to perform a variety of desirable tasks in statistical learning. 
	\item Dimension reduction: under the hypothesis that the data lie on, or are concentrated near, a manifold $\mathcal{M} \subset \Reals^d$ of dimension $m < d$, the eigenvectors $v_1,\ldots,v_m$, can be used to embed $x_i \mapsto (v_{i,1},\ldots v_{i,m})$, reducing the dimension from $d$ to $m$ without losing the geometric structure of $\Pbb$. This algorithm is known as Laplacian Eigenmaps (Belkin).
	\item \textcolor{red}{Clustering: under the hypothesis that the distribution $\Pbb$ is supported on disjoint connected components $\mathcal{C}_1,\ldots,\mathcal{C}_k$, the span of the smallest $k$ eigenvectors of the Laplacian operator $L_n$ will approximately equal the span of the functions $\phi_j(x) = \1\set{x \in \mathcal{C}_j}, j = 1,\ldots,k$.} \textcolor{red}{(Belkin + Arias-Castro)}
	\item Semi-supervised and supervised learning: Discuss the study of Laplacians in supervised learning \textcolor{red}{(Bousquet + Chapelle, Lee + Izbicki, GarciaTrillos + Murray, Padilla)} Discuss the study of Laplacians in semi-supervised learning \textcolor{red}{(Belkin + Niyogi, Zhou + Belkin, Slepcev and Thorpe)}. 
\end{itemize}

In this thesis proposal, we pick up a few of these threads, showing how spectral methods can be used to perform the statistical tasks of density clustering and nonparametric hypothesis testing. Before doing so, we provide some background on each of these areas.

\subsection{Clustering.}

Clustering involves splitting a given data set
into groups that satisfy some notion of within-group similarity and
between-group difference. There are many ways to define different types of clusters; we review a few that are especially related to our work.

\subsubsection{Density clustering.}

Let $X = \set{x_1,\ldots,x_n}$ consist of i.i.d. samples drawn from a distribution $\Pbb$ on $\Reals^d$, with density $f$. The \emph{density clusters} of $f$, defined as the connected components of
the upper level set $\{x \in \Rd : f(x) \geq \lambda\}$ for some $\lambda > 0$
are a central object of interest in the statistical clustering literature, dating
back to the work of \citet{hartigan1981}, who shows that the classical single-linkage hierarchical clustering algorithm is guaranteed to recover density clusters only when $d = 1$. \citet{polonik1995, rigollet2009} study density clustering under the 
symmetric set difference metric, \citet{tsybakov1997, singh2009} describe
minimax optimal level-set estimators under Hausdorff loss and
\citet{hartigan1981, chaudhuri2010, balakrishnan2013,kpotufe11} consider consistent estimation of the
cluster tree. None of these works, however, study the ability of spectral algorithms to recover density clusters, which will be our focus.

\subsubsection{Nonparametric mixture model.}

The results in Section~\ref{subsec:neighborhood_graphs_spectra} detailing with spectral convergence of the graph Laplacian operator have straightforward consequences for the consistency of spectral clustering (for instance, see \cite{vonluxburg2008,garciatrillos18}). However, the behavior of the spectra of these continuum operators can in general be hard to grasp. Therefore, further work relating this spectra to the geometry of the distribution $\Pbb$ is of interest. In this spirit, \citet{shi2009,schiebinger2015,garciatrillos19,ariascastro11} examine the ability of spectral algorithms to recover the latent labels in certain geometrically well-conditioned nonparametric mixture models $\Pbb = \sum_{i = 1}^{k} \pi_i \Pbb_i$. These results focus on global rather than local 
methods, and thus impose global rather than local conditions on the nature
of the density. Moreover, they do not in general guarantee recovery of density 
clusters, which is the focus in our work. Perhaps most importantly, these works
rely on general cluster saliency conditions, which implicitly depend on many
distinct geometric aspects of the cluster $\Cset$ under consideration. We will make
this dependence more explicit, and in doing so expose the role each geometric
condition plays in the clustering problem.

\subsubsection{Worst-case analysis.}
\label{subsubsec:worst_case_analysis}

An altogether different approach is to treat the graph $G = (V,E)$ as fixed, rather than as a random geometric graph formed over data as is the case when $G = G_{n,r}$. Now, the quality of the clustering algorithm necessarily depends on the extent to which $G$ contains natural clusters. To quantify the \textcolor{red}{clusterability} of a given subset $S \subseteq V$, let the normalized cut $\Phi(S;G)$ be given by
\begin{equation*}
\Phi(S;G) = \frac{\cut(S;G)}{\min\{\vol(S;G), \vol(S^c;G)\}},\quad \cut(S;G) := \sum_{u \in S} \sum_{w \in S^c} A_{uv}, \quad \vol(S;G) = \sum_{u \in S} D_{uu}.
\end{equation*}
and the conductance $\Psi(S;G)$ be 
\begin{equation*}
\Psi(S;G) = \min_{A \subset S} \Phi(A;G[S])
\end{equation*}
where $G[S] = (S,E_S), E_S = E \cap (S \times S)$ is the subgraph of $G$ induced by $S$. The normalized cut and conductance functionals measure the external connectivity of $S$ to the rest of $G$, and the internal connectivity of $G[S]$, respectively. \cite{kannan04} assume there exists a partition $S_1 \cup \ldots \cup S_k$ such that $\Phi(S_j) \leq \phi$ and $\Psi(S_j) \geq \psi$ for all $j$. They analyze an algorithm which recursively splits the graph $G$ based on its eigenvectors, resulting in clusters $C_1 \cup \ldots \cup C_k = V$, and prove upper bounds on the normalized cuts $\Phi(C_j)$ and conductances $\Psi(C_j)$ in terms of $\phi$ and $\psi$. \citet{spielman2013,andersen2006} consider the local clustering problem, where assumptions are made only with respect to some small subset $S \subseteq V$. They each consider random walk based algorithms which cluster using locally-biased spectra, and derive upper bounds on the normalized cut of the outputted cluster $\Phi(C)$ in terms of $\Phi(S)$. \cite{zhu2013} extends this analysis, showing that improvements in the quality of a cluster estimate are possible when $\Psi(S) > \Phi(S)$. We rely on the analysis of \cite{zhu2013} in our work; however, our setting is quite different, as we work with respect to the random graph $G_{n,r}$ rather than a fixed $G$, and seek to derive bounds which depend on the distribution $\Pbb$ rather than on $\Phi$ and $\Psi$, which in our context are random functionals.

\subsection{Nonparametric hypothesis testing.}

In hypothesis testing, the goal is distinguish whether our data are generated from some null model -- generally a simple model displaying a lack of meaningful structure -- or an alternative model. \emph{Regression} and \emph{density} testing are two related but distinct hypothesis testing problems. We will formally introduce each problem and provide some relevant background. Additionally, we will cover some common classes of test statistics used for nonparametric hypothesis testing.

\subsubsection{Regression testing.}

In addition to observing i.i.d samples $x_1,\ldots,x_n$ from a distribution $\mathbb{P}$ with support $\mathcal{X} := [0,1]^d \subset \Reals^d$\footnote{For the purposes of this proposal, whenever we discuss hypothesis testing we will assume $\mathcal{X} = [0,1]^d$, i.e. the unit cube with dimension equal to that of the ambient space $\Reals^d$.}, suppose we observe $Y = \set{y_1,\ldots,y_n}$ according to the following regression model:
\begin{equation}
\label{eqn:regression}
y_i = f(x_i) + \varepsilon_i, \quad \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
\end{equation}
We wish to distinguish
\begin{equation*}
\mathbf{H}_0: f = f_0 := 0 \quad \textrm{vs} \quad \mathbf{H}_{\textrm{a}}: f \neq f_0
\end{equation*}
We will evaluate our performance using worst-case risk: for a given function class $\mathcal{H}$ and test function $\phi: \Reals^n \to \set{0,1}$, let
\begin{equation*}
\mathcal{R}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, f \neq f_0} \Ebb_f(1 - \phi).
\end{equation*}
Without further assumptions on $\mathcal{H}$ no test achieves meaningful power uniformly over all $f \in \mathcal{H}$ \citet{janssen2000}; to make the hypothesis testing question well-posed we will therefore restrict the function class $\mathcal{H}$ in two standard ways. One, when considering type II error we will remove a radius-$\epsilon$ ball in the $L_2$ norm around $f_0$ and consider only the remaining functions
\begin{equation*}
\mathcal{H}_{\epsilon} := \set{f \in \mathcal{H}:\norm{f - f_0}_2 > \epsilon}.
\end{equation*}
Otherwise we can always pick a function $f \neq f_0$ in $\mathcal{H}$ which is indistinguishable from $f_0$ given a finite number of samples. Two, we will insist that the functions in $\mathcal{H}$ display some degree of regularity or smoothness; otherwise it may not be possible to consistently estimate the functional $\norm{f - f_0}_2$.

Let $\alpha \in (0,1)$ be a user input specifying a tolerated level of testing error. A typical way to characterize the performance of a test $\phi$ is through the critical radius $\epsilon_n(\phi; \mathcal{H})$, defined as
\begin{equation*}
\epsilon_n(\phi; \mathcal{H}) = \inf\{\epsilon: \mathcal{R}(\phi; \mathcal{H}_{\epsilon}) \leq \alpha\}.
\end{equation*}
We can then measure the difficulty of a given problem through the minimax critical radius $\epsilon_n^{\star}$, formally 
\begin{equation*}
\epsilon_n^{\star}(\mathcal{H}) = \inf \set{\epsilon:\inf_{\phi} \mathcal{R}(\phi; \mathcal{H}_{\epsilon} ) \leq \alpha}.
\end{equation*}
where the infimum is over all Borel measurable functions $\phi: \Reals^n \to \set{0,1}$. We say a test $\phi$ is minimax optimal over a function class $\mathcal{H}$ if $\epsilon_n(\phi; \mathcal{H}) \asymp \epsilon_n^{\star}(\mathcal{H})$, and we call the rate at which $\epsilon_n^{\star}(\mathcal{H}) \to 0$ as a function of $n$ the \emph{minimax testing rate} over $\mathcal{H}$. To concisely state rates, we will use order notation, in which we write $a_n = O(b_n)$ when $a_n \leq c b_n$ for some constant $c$ and all $n \in \mathbb{N}$, $a_n = o(b_n)$ when $a_n/b_n \to 0$ as $n \to \infty$, and $a_n \asymp b_n$ when $c b_n \leq a_n \leq C b_n$ for constants $c$ and $C$ and all $n \in \mathbb{N}$.

\paragraph{Testing over Sobolev classes.}
The first type of function class $\mathcal{H}$ we consider will be balls in Sobolev spaces. We say a Borel measurable function $f: \mathcal{X} \to \Reals$ is in the space $\mathcal{L}^p(\mathcal{X})$ for $1 \leq p < \infty$ if $\int_{\mathcal{X}} \abs{f(x)}^p \,dx < \infty$. Then, the Sobolev space $W_d^{s,2}(\mathcal{X})$ consists of all functions $f \in \mathcal{L}^1(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$(See \citet{evans10} for a formal definition) belongs to $\mathcal{L}^p(\mathcal{X})$. The Sobolev $\{s,2\}$ norm is then 
\begin{equation*}
\norm{f}_{W_d^{s,2}(\mathcal{X})}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^2 \,dx
\end{equation*}
and for a given $L > 0$, the corresponding ball is $W_d^{s,2}(\Xset; L) = \set{f: \norm{f}_{W^{s,2}(\Xset)} \leq L}$. In this work, we will treat $L$ as a fixed, known constant.

Minimax testing rates for the Sobolev regression testing problem in one dimension were first derived by \citet{ingster82} for the idealized \emph{white noise} regression model:
\begin{equation}
\label{eqn:gaussian_white_noise}
\,dY(x) = f(x) \,dx + \frac{\sigma}{\sqrt{n}}\,dW(x)
\end{equation}
where $\,dW(x)$ denotes a Gaussian white noise process on $\mathcal{X} \subset \Reals$ (defined formally in \citet{gine16} Section 1.2) and $\sigma > 0$ is a known constant. In this case, the minimax test rate is $\epsilon_n^{\star}(\mathcal{W}_d^{s,2}(\mathcal{X}, L)) \asymp n^{-2s/(4s + 1)}$ for all values of $s > 0$. \citet{ingster05} extend this to the multivariate setting where $\mathcal{X} \subseteq \Reals^d$, and show that $\epsilon_n^{\star}(\mathcal{W}_d^{s,2}(\mathcal{X}, L)) \asymp n^{-2s/(4s + d)}$, again for all $d \geq 1, s > 0$. The case where we observe data according to the generative model~\eqref{eqn:regression} was treated in \cite{ingster09}, wherein it is shown that the minimax rate remains the same  $\epsilon_n^{\star}(\mathcal{W}_d^{s,2}(\mathcal{X}, L)) \asymp n^{-2s/(4s + d)}$ so long as $4s > d$. When $4s < d$, if we replace the assumption $f \in W_d^{s,2}(\mathcal{X},L)$ by the assumption $\left(\int_{\mathcal{X}} f^4(x) \,dx\right)^{1/4}$, then \citet{guerre02} show that a test $\phi$ based on the statistic $\frac{1}{n}\sum_{i = 1}^{n}y_i^2$ will have testing error on the order of $n^{-1/4})$ . Since $\mathcal{W}^{s,2}$ does not compactly embed into $\mathcal{L}^4$ when $4s < d$ this result does not apply to Sobolev spaces in this regime; indeed it seems likely that uniformly consistent testing over these spaces is not possible, as the function $f^2$ may no longer have finite variance.

\subsubsection{Density testing.}
\label{subsubsec:density_testing}

In the two-sample density testing problem, we observe independent samples $Z = z_1,\ldots,z_N \sim \Pbb$ and $Y = y_1,\ldots,y_M \sim \Qbb$, where $\Pbb$ and $\Qbb$ are distributions over $\Reals^d$ with densities $p$ and $q$, respectively. Our goal is to distinguish the hypotheses
\begin{equation*}
\mathbf{H}_0: \Pbb = \Qbb \quad \textrm{vs} \quad \mathbf{H}_{\textrm{a}}: \Pbb \neq \Qbb
\end{equation*}
and we again evaluate our performance using worst-case risk; letting $\phi:\Reals^{N + M} \to \{0,1\}$, 
\begin{equation*}
\mathcal{R}(\phi; \mathcal{H}) = \inf_{p \in \mathcal{H}}\Ebb_{p,p}(\phi) + \sup_{p \neq q, p,q \in \mathcal{H}} \Ebb_{p,q}(1 - \phi),
\end{equation*}
and the critical radius $\epsilon_n$. (Although we overload notation by having $\mathcal{R}$ and $\epsilon_n$ refer to the worst-case risk and critical radius in both the regression and density testing problems, it will always be clear from context which problem we are referring to.)
\paragraph{Density testing in Holder spaces.} Unlike in the regression testing context, where $f$ is assumed to belong to a Sobolev space, most of the study of density testing seems to have dealt with the case where $\mathcal{H}$ is Holder space. For a given $s > 0$, the $s$th Holder norm is given by
\begin{equation*}
\norm{f}_{C_d^{s}(\mathcal{X})} := \sum_{\abs{\alpha} \leq s} \norm{D^{\alpha}f}_{\infty} + \sum_{\abs{\alpha} = s} \sup_{x,y \in \mathcal{X}} \frac{\abs{D^{\alpha}f(y) - D^{\alpha}f(x)}}{\norm{x - y}_2}
\end{equation*}
and the $s$th Holder space $C_d^{s}(\mathcal{X})$ consists of all functions which are $s$ times continuously differentiable with finite $s$ Holder norm. Denote the Holder unit ball by $C_d^{s}(\mathcal{X},L) = \set{f \in C_d^{s}(\mathcal{X}): \norm{f}_{C_d^{s}(\mathcal{X})} \leq L}$.

\citet{ingster87} considers the 1d version of the problem and shows that the minimax critical radius for the Holder balls $C_d^{s}(\mathcal{X};L)$ is $\epsilon_n^{\star}(C^{s}(\mathcal{X};L)) \asymp n^{-2s/(4s + 1)}$. This means that the Holder space $C_d^{s}(\mathcal{X})$ and the smallest Sobolev space it compactly embeds into, $W_d^{s,2}(\mathcal{X})$, have the same minimax testing rates in the density and regression testing problems, respectively. \cite{ariascastro18} examines the multivariate version of the problem and derives the rate $\epsilon_n^{\star}(C_d^{s}(\mathcal{X};L)) \asymp n^{-2s/(4s + d)}$. In both cases, the stated rates hold for all $s > 0$, and in the second case for all $d \geq 1$.

Using asymptotic equivalence results \citet{reiss08,nussbaum96,brown96} between the density and regression setups, and sampling and Gaussian white noise models, we can infer some results about minimax rates over Sobolev classes in the former from those in the latter. However, to the best of our knowledge, the density testing problem over Sobolev spaces has not been directly studied.

\subsubsection{Background on nonparametric test statistics.}

The $\chi^2$-test is a classical two-sample test, in which the domain $\mathcal{X}$ is partitioned into bins $B_1,\ldots,B_\kappa$ where $\kappa \in \mathbb{N}$ is a tuning parameter. Letting $\mathbb{P}_N$ be the empirical distribution of $Z$, and likewise for $\mathbb{Q}_M$ and $Y$, the resulting statistic can be expressed as 
\begin{equation}
\label{eqn:chi_squared}
T_{\chi^2} := \sum_{k = 1}^{\kappa} \left(\frac{\mathbb{P}_N(B_k) - \mathbb{Q}_M(B_k)}{\mathbb{P}_N(B_k) + \mathbb{Q}_M(B_k)}\right)^2.
\end{equation}
The upper bounds in the density testing problem discussed in the previous section are derived using an unnormalized version of $T_{\chi^2}$. In addition to the $\chi^2$-test, there are several classes of test statistics used to perform nonparametric hypothesis testing which are worthy of our attention.

\paragraph{Edge-counting.}
In the two-sample density testing problem, \citet{schilling86, bhattacharya15} analyze edge-counting statistics, where a $k$NN graph $G_{n,k}$ is formed over all the samples $X = (z_1,\ldots,z_N,y_1,\ldots,y_M)$ (where $n = M + N$ is the total number of samples) and the statistic considered is then
\begin{equation}
\label{eqn:edge_count_statistic}
T_{\textrm{edge}} := \sum_{i = 1}^{N} \sum_{j = 1}^{M} \1(z_i \sim y_j \textrm{ in } G_{n,k}).
\end{equation}
While there are some similarities between $T_{\mathrm{edge}}$ and the statistic $T_{\mathrm{spec}}$ we propose and analyze later in Section~\ref{subsec:spectral_regression_testing}, they are quite distinct. For one, they depend on different constructions of the neighborhood graph; perhaps more fundamentally, they use the spectrum of the resulting neighborhood graphs in different ways. Preliminary results (see Figure~\ref{fig:fig3}) show that these differences in construction lead to empirically different behavior over a range of testing problems.

\paragraph{Truncated-series.}

In the regression testing problem, \citet{ingster09} upper bound the minimax testing rate by considering a truncated series test statistic, where the data is projected onto a subspace spanned by (evaluations of) low-frequency harmonics, and the test statistic is the empirical norm of this projection. Let $\set{\phi_j}$ be the standard Fourier basis in $L_2([0,1])$, meaning
\begin{equation*}
\phi_0(x) = 1,~~ \phi_j(x) = \sqrt{2}\cos(2\pi j u) ~~\textrm{for each $j \in \mathbb{Z}_{-}$},~~ \phi_j(x) = -\sqrt{2}\sin(2\pi j u) ~~\textrm{for each $j \in \mathbb{Z}_{+}$}:
\end{equation*}
The test statistic which \citet{ingster09} analyze is then
\begin{equation}
\label{eqn:harmonic}
T_{\mathrm{Harm}} := \frac{1}{n}\sum_{\ell,-\ell \in [\kappa]^d} \left(\sum_{i = 1}^{n} \phi_{\ell}(x_i) y_i\right)^2, \quad \textrm{where}~ \phi_{\ell}(x) = \prod_{i = 1}^{d} \phi_{\ell_i}(x_i)
\end{equation}
for $\ell \in \mathbb{Z}^d$, $x \in \Reals^d$, and 
$\kappa \in \mathbb{N}$ a tuning parameter. The test statistic $T_{\mathrm{spec}}$ which we propose and analyze in Section~\ref{subsec:spectral_regression_testing} is closely related to $T_{\mathrm{Harm}}$, the crucial difference being that instead of the Fourier basis $\set{\phi_{\ell}}$, which are eigenfunctions of the Laplacian operator over $[0,1]^d$, we use eigenvectors $\set{v_k}$ of the graph Laplacian matrix.

\paragraph{Integral Probability Metrics (IPMs).}

Much recent attention \citet{muller97,sriperumbudur09,gretton12,mroueh17} has been paid to IPMs, which are statistics of the form
\begin{equation*}
\sup_{f \in \mathcal{F}} \abs{\mathbb{P}_n f - \mathbb{Q}_n f}
\end{equation*}
used to perform two-sample density testing. (Here, $\mathcal{F}$ is appropriately chosen set of functions, for example $\mathcal{F}$ might be the unit ball in a Holder or Sobolev space). Of course, for any graph $G = (V,E)$, the corresponding incidence matrix $B$ induces a notion of smoothness for vectors $\theta \in \Reals^{\abs{V}}$. Formally for some $s,p \in \mathbb{N}$, let the smoothness functional $S_{s,p}(\theta) = \norm{B^{(s)}\theta}_p$ where
\begin{equation*}
B^{(s)} = 
\begin{cases}
L^{s/2}, & \textrm{when $s$ is even} \\
B L^{(s-1)/2}, & \textrm{when $s$ is odd.}
\end{cases}
\end{equation*}
Then the classes $\Theta_{s,p} := \set{\theta \in \Reals^n: S_{s,p}(\theta) \leq L}$ can be viewed as possessing various types of smoothness, analogous to function classes such as Sobolev or Holder classes. It is therefore natural to consider IPMs of the form
\begin{equation*}
\sup_{\theta \in \Theta_{s,p}} \abs{\mathbb{P}_N \theta - \mathbb{Q}_M \theta};
\end{equation*}
and we will propose a pair of such statistics in Section~\ref{sec:neighborhood_graph_IPMs}.

\section{Density clustering with PPR.}

In our density clustering work, we examine the ability of Personalized PageRank (PPR) -- a local spectral algorithm-- to recover density clusters. First, we motivate and formally define the PPR clustering algorithm we subsequently analyze.

\subsection{PPR clustering.}

When applied to geometric graphs built from a large number of samples,
global spectral clustering methods can be computationally cumbersome and   
insensitive to the local geometry of the underlying distribution
\citep{leskovec2010,mahoney2012}.  This has led to increased interest in
\emph{local} spectral clustering algorithms, which leverage locally-biased spectra
computed using random walks around some user-specified seed node.  A popular 
local clustering algorithm is the Personalized PageRank (PPR) algorithm, first introduced by 
\citet{haveliwala2003}, then further developed by
several others~\citep{spielman2011,spielman2014,andersen2006,mahoney2012,zhu2013}.

The PPR vector $p_v = p(v,\alpha;G_{n,r})$, based on
a seed node $v \in V$ and a teleportation parameter $\alpha \in [0,1]$, is defined to be 
the solution of the following linear system:
\begin{equation}
\label{eqn: ppr_vector}
p_v = \alpha e_{v} + (1 - \alpha) p W,
\end{equation}
where $W = \frac{1}{2}(I + D^{-1}A)$ is the lazy random walk matrix over
$G_{n,r}$ and $e_{v}$ is the indicator vector for node $v$ (that has a 1 in the
$v$-th position and 0 elsewhere).

\subsubsection{Sweep cut.}

We employ the sweep cut method to obtain a local cluster from the PPR embedding $p_v$. In more detail, for a level $\beta > 0$, we define a \emph{$\beta$-sweep cut} of $p_v = (p_v(u))_{u \in V}$
as   
\begin{equation}
\label{eqn: sweep_cuts}
S_{\beta,v} := \set{u \in V: \frac{p_v(u)}{D_{uu}} > \beta}.
\end{equation}
We will use the normalized cut metric to determine which sweep cut $S_{\beta}$
is the best cluster estimate. For a set $S \subseteq V$ with complement $S^c = V
\setminus S$, we define \smash{$\cut(S;G_{n,r}) := \sum_{u \in S, v \in S^c}
	A_{uv}$}, and \smash{$\vol(S; G_{n,r}) := \sum_{u \in S} D_{uu}$}.  We
define the \emph{normalized cut} of $S$ as
\begin{equation}
\label{eqn: normalized_cut}
\Phi(S; G_{n,r}) := \frac{\cut(S;G_{n,r})}{\min \set{\vol(S; G_{n,r}), \vol(S^c; G_{n,r})}}.
\end{equation}
Having computed sweep cuts $S_{\beta}$ over a range \smash{$\beta \in  
	(L, U)$} (where the range $(L,U)$ is also a user-specified parameter), we output the cluster estimate
\smash{$\widehat{C} = S_{\beta^*}$} that has minimum normalized cut.
%\smash{$\Phi(S_{\beta^*}; G_{n,r})$}. 
For concreteness, the PPR algorithm is summarized in Algorithm~\ref{alg: ppr}.

\begin{algorithm}
	\caption{PPR on a neighborhood graph}
	\label{alg: ppr}	
	{\bfseries Input:} data $X=\{x_1,\ldots,x_n\}$, radius $r > 0$, teleportation
	parameter $\alpha \in [0,1]$, seed $v \in X$, sweep cut range $(L,U)$. \\     
	{\bfseries Output:} cluster $\widehat{C} \subseteq V$.
	\begin{algorithmic}[1]
		\STATE Form the neighborhood graph $G_{n,r}$.
		\STATE Compute the PPR vector $p_v=p(v, \alpha; G_{n,r})$ as in \eqref{eqn: 
			ppr_vector}. 
		\STATE For \smash{$\beta \in (L,U)$} compute sweep cuts 
		$S_{\beta}$ as in \eqref{eqn: sweep_cuts}.
		\STATE Return the cluster \smash{$\widehat{C} = S_{\beta^*}$}, where  
		$$
		\beta^* = \argmin_{\beta \in (L,U)}~ \Phi(S_{\beta}; G_{n,r}).
		$$
	\end{algorithmic}
\end{algorithm}

\subsection{Estimation of density clusters} 
Let \smash{$\Cbb_f(\lambda)$} denote 
the connected components of the density upper level set $\{x \in \Rd: f(x) >
\lambda\}$.  For a given density cluster \smash{$\Cset \in \Cbb_f(\lambda)$}, we
call $\Cset[X] = \Cset \cap X$ the \emph{empirical density cluster}. The size of the symmetric set difference between estimated and empirical cluster is 
a commonly used metric to quantify cluster estimation error
\citep{korostelev1993,polonik1995,rigollet2009}. We will consider a related metric, the volume of the symmetric set difference, which weights points according to their degree in $G_{n,r}$. 
\begin{definition}
	\label{def:volume_symmetric_set_difference}
	For an estimator $\widehat{C} \subseteq X$ and a set $\mathcal{S} \subseteq \Rd$, the symmetric set difference is 
	\begin{equation*}
	\widehat{C} \vartriangle \mathcal{S}[X] := \bigl(\widehat{C} \setminus \mathcal{S}[X]\bigr) \cup \bigl(\mathcal{S}[X]\setminus\widehat{C} \bigr).
	\end{equation*}
	We will measure the discrepancy between $\widehat{C}$ and $\mathcal{S}[X]$ by the volume of the symmetric set difference,
	\begin{equation*}
	\Delta(\widehat{C}, \mathcal{S}[X]) := \vol(\widehat{C} \vartriangle \mathcal{S}[X]; G_{n,r})
	\end{equation*}
\end{definition}
However, the symmetric set difference does not measure whether \smash{$\widehat{C}$} 
can distinguish any two distinct clusters \smash{$\Cset,\Cset' \in
	\Cbb_f(\lambda)$}. We therefore also study a second notion of cluster
estimation, first introduced by \citet{hartigan1981}, and defined
asymptotically.

\begin{definition}[Consistent density cluster estimation]
	\label{def: consistent_density_cluster_estimation}
	For an estimator \smash{$\widehat{C} \subseteq X$} and cluster 
	\smash{$\Cset \in \Cbb_f(\lambda)$}, we say \smash{$\widehat{C}$} is a consistent 
	estimator of $\Cset$ if for all \smash{$\Cset' \in \Cbb_f(\lambda)$} with
	$\Cset \not= \Cset'$, the following holds as $n \to \infty$: 
	\begin{equation}
	\label{eqn: consistent_density_cluster_recovery}
	\Cset[X] \subseteq \widehat{C} \quad \text{and} \quad
	\widehat{C} \cap \Cset'[X] = \emptyset,
	\end{equation}
	with probability tending to 1.
\end{definition}
\noindent Consistent cluster recovery roughly ensures that, for a given density threshold $\lambda$, the estimated cluster $\widehat{C}$ contains all
points in a true
density cluster $\Cset \in \Cbb_f(\lambda)$, and simultaneously does not contain any points in any other density cluster $\Cset' \in \Cbb_f(\lambda)$.

With these definitions in place, our broad goal will be to understand the extent to which the PPR algorithm is able to
recover a cluster which either guarantees low symmetric set difference to a true density cluster, or which consistently estimates
a true density cluster.

\subsection{Geometric Conditioning of a Density Cluster.}%: Geometric Conditions and the Condition Number of a Density Cluster}
%We formalize some geometric conditions, and use these to define a condition
%number $\kappa(\Cset)$, which measures the difficulty PPR will have in  
%estimating $\Cset$. (Our theoretical guarantees for PPR will be framed in terms 
%of $\kappa(\Cset)$.)
%\paragraph{Geometric conditions on density clusters.} 
At a high level, for PPR
to be successful, the underlying density cluster must be geometrically
well-conditioned.  A basic requirement is that we need to avoid clusters which contain arbitrarily
thin bridges or spikes. As in the work of \citet{chaudhuri2010}, we consider a
thickened version of \smash{$\Cset \in \Cbb_f(\lambda)$} defined as 
$\Csig := \set{x \in \Reals^d: \dist(x,\Cset) \leq \sigma}$, which 
we call the \emph{$\sigma$-expansion} of $\Cset$. Here 
\smash{$\dist(x,\Cset) := \inf_{y \in \Cset} \norm{y - x}$}.  We now list our
conditions on $\Csig$.

\begin{enumerate}[label=(A\arabic*)]
	\item
	\label{asmp: bounded_density}
	\emph{Bounded density within cluster:} There exist constants
	$0<\lambda_{\sigma}< \Lambda_{\sigma}<\infty$ such that 
	$\lambda_{\sigma} \leq \inf_{x \in \Csig} f(x) \leq \sup_{x \in \Csig} f(x)
	\leq \Lambda_{\sigma}$.

	\item 
	\label{asmp: low_noise_density}
	\emph{Low noise density:} There exists $c_0 > 0$ and $\gamma \in [0,1]$ such that for any $x \in \Rd$ with $0 < \dist(x, \Csig) \leq \sigma$,   
	\begin{align*}
	\inf_{x' \in \Csig} f(x') - f(x) \geq  c_0 \dist(x, \Csig)^{\gamma}. 
	\end{align*}
	Roughly, 
	this assumption ensures that the density decays sufficiently quickly as we move away from
	the target cluster $\Csig$, and is a standard assumption in the level-set estimation literature
	(see for instance~\cite{singh2009}).

	\item
	\label{asmp: embedding}
	\emph{Lipschitz embedding:}
	There exists $g: \Reals^d \to \Reals^d$ with the following properties: 
	\begin{enumerate}
		\item we have $\Csig = g(\mathcal{K})$, for a convex set $\mathcal{K} \subseteq \Rd$
		with $\mathrm{diam}(\mathcal{K}) = \sup_{x,y \in \mathcal{K}}\norm{x - y} =:
		\diam < \infty$
		\item $\det(\nabla g (x)) = 1$ for all $x \in \Csig$, where
		$\nabla g(x)$ is the Jacobian of $g$ evaluated at $x$ and
		\item for some $L
		\geq 1$,   
		\begin{equation*}
		\norm{g(x) - g(y)} \leq L \norm{x - y} ~
		\text{for all $x,y \in \mathcal{K}$}. 
		\end{equation*}
	\end{enumerate}
	Succinctly, we assume that $\Csig$ is the image of a convex set with finite diameter 
	under a  measure preserving, Lipschitz transformation.

	\item
	\label{asmp: bounded_volume}
	\emph{Bounded volume:}
	For a set $\mathcal{S} \subseteq \Reals^d$, define the $\Pbb$-weighted volume of $\mathcal{S}$ to be
	\begin{equation}
	\label{eqn:volume}
	\vol_{\Pbb,r}(\mathcal{S}) := \int_{\mathcal{S}} \Pbb(B(x,r)) \,d\Pbb(x).
	\end{equation}
	where $B(x,r)$ is the closed ball of radius $r$ centered at $x$.

	We assume that the neighborhood graph radius $0 < r \leq \sigma/2d$ is chosen such that
	\begin{equation*}
	\vol_{\Pbb,r}(\mathcal{S}) \leq \frac{1}{2}\vol_{\Pbb,r}(\Reals^d).
	\end{equation*}
\end{enumerate}

\subsubsection{Condition number} 
We now introduce the condition number $\kappa(\Cset)$. As we will see in Theorems~\ref{thm:volume_ssd} and ~\ref{thm: consistent_recovery_of_density_clusters},
the smaller $\kappa(\Cset)$ is, the
more success PPR will have in recovering the target cluster $\Cset$.

\begin{definition}[Well-conditioned density clusters]
	\label{def:well_conditioned_density_cluster}
	For $\lambda > 0$ and \smash{$\Cset \in \Cbb_f(\lambda)$}, let $\Cset$ satisfy 
	\ref{asmp: bounded_density}--\ref{asmp: bounded_volume} for some $\sigma > 0$. Then, for universal constants $c_1, c_2, c_3 > 0$ to be specified
	later, we set 
	\begin{equation}
	\label{eqn: condition_number_1}
	\Phi_u(\Cset) 
	:= c_1 r \frac{d}{\sigma} \frac{\lambda}{\lambda_{\sigma}}
	\frac{(\lambda_{\sigma} - c_0 \frac{r^{\gamma}}{\gamma +
			1})}{\lambda_{\sigma}},~  
	\tau_u(\Cset) := c_2 \frac{\Lambda_{\sigma}^4 d^3 \rho^2
		L^2}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{\Lambda_{\sigma}}{\lambda_{\sigma}^2r}\right) + c_3,
	\end{equation}
	and letting \smash{$\kappa(\Cset) := \Phi_{u}(\Cset) \cdot
		\tau_{u}(\Cset)$}, we call $\Cset$ a 
	\emph{$\kappa$-well-conditioned density cluster}.  
\end{definition}

\textbf{Remark:}~~ The population-level quantities $\Phi_u(\Cset)$ and  $\tau_{u}(\Cset)$ measure the internal and external connectivity, respectively, of the thickened density cluster $\Csig$. In particular, as we will see in Section~\ref{sec: analysis}, we will derive that $\Phi_u(\Cset)$ is an upper bound on the normalized cut $\Phi(\Csig[X];G_{n,r})$, and that $\tau_{u}(\Cset)$ is an upper bound on the mixing time $\tau_{\infty}(G_{n,r}[\Csig[X]])$ (defined later in~\eqref{eqn: mixing_time}). These bounds on graph functionals translate to estimates of PPR cluster quality via the results we mentioned in Section~\ref{subsubsec:worst_case_analysis}.

\subsubsection{Well-initialized algorithm} As is typical in the local clustering
literature, our algorithmic results will be stated with respect to specific
ranges of each of the user-specified parameters. In particular, for a
well-conditioned density cluster $\Cset$, we require that some of the tuning parameters of Algorithm~\ref{alg: ppr} are chosen to fall within specific ranges,
\begin{equation}
\begin{gathered}
\label{eqn: initialization}
0 < r \leq \frac{\sigma}{2d}, \quad 
\alpha \in [1/10, 1/9) \cdot \frac{1}{\tau_u(\theta)}, \\
(L,U) \subseteq (1/50,1/5) \cdot \frac{1}{2{n \choose 2}\vol_{\Pbb,r}(\Csig)}.
\end{gathered}
\end{equation}

\begin{definition}
	If the input parameters to Algorithm \ref{alg: ppr} satisfy \eqref{eqn:
		initialization} for some well-conditioned density cluster $\Cset$, we
	say the algorithm is \emph{well-initialized}. 
\end{definition}
\noindent In practice it is  not feasible to set hyperparameters based on the
underlying (unknown) density $f$. Typically, one tunes PPR over a range of
hyperparameters and selects the cluster which has the smallest normalized cut.

\subsection{Cluster Recovery in Symmetric Set Difference}
We now present a bound on the volume of the symmetric set difference between the estimated cluster $\widehat{C}$ and empirical cluster $\Csig[X]$. 
In this theorem, and hereafter, we let 
$c, c_i > 0$ represent universal constants, and 
$c(\Pbb,r), c_i(\Pbb,r) > 0$ represent constants which may depend on $\Pbb,\lambda$ and $r$ but not on the sample size $n$; where possible, we keep track of all constants in our proofs. This bound is stated with respect to the volume $\vol(\Csig[X];G_{n,r})$, which we will shorten to $\vol_{n,r}(\Csig[X])$ to keep things simple.
\begin{theorem}
	\label{thm:volume_ssd}
	Fix $\lambda > 0$ and let \smash{$\Cset \in \Cbb_f(\lambda)$} be a
	$\kappa$-well-conditioned density cluster for some $\sigma > 0$. If Algorithm \ref{alg: ppr} is \emph{well-initialized} with respect to $\Cset$, then 
	%there exists a universal constant $c > 0$, constants in sample size $c_1(\Pbb,r)$ - $c_6(\Pbb,r)$, and 
	there exists a set $\Csig[X]^g \subseteq \Csig[X]$ of large volume 
	with $\vol_{n,r}(\Csig[X]^g) \geq \frac{1}{2}\vol_{n,r}(\Csig[X])$ such that the following statement holds: let Algorithm~\ref{alg: ppr} be run with any seed node $v \in \Csig[X]^g$. Then, for any $n$ large enough so that
	\begin{equation*}
	n \geq c_1(\Pbb,r)\cdot (\log n)^{\max\{\frac{3}{d},1\}}
	\end{equation*}
	the volume of the symmetric set difference is upper bounded
	\begin{equation}
	\label{eqn:volume_ssd}
	\Delta(\Csig[X], \widehat{C}) \leq c \cdot \kappa(\Cset) \cdot \vol_{n,r}(\Csig[X]),
	\end{equation}
	with probability at least $1 - \frac{c_2(\Pbb,r)}{n}$. 
\end{theorem}
This result establishes that the volume of the symmetric set difference \smash{$\Delta(\Csig[X],
	\widehat{C})$} is upper-bounded by a quantity proportional to the difficulty of the clustering problem, as
measured by the condition number $\kappa(\Cset)$. The primary technical difficulty involved in proving Theorem~\ref{thm:volume_ssd} comes from upper bounding the random graph functionals $\Phi(\Csig[X];G_{n,r})$ and $\tau_{\infty}(G_{n,r}[\Csig[X]])$; once this is accomplished, the result is a straightforward implication of \citet{zhu2013}.

\subsection{Consistent Cluster Recovery}

Theorem \ref{thm: consistent_recovery_of_density_clusters} implies that when $\kappa(\Cset)$ is sufficiently small, \smash{$\widehat{C}$} will be a consistent  
density cluster estimator. We will need one additional regularity condition, to preclude arbitrarily low degree
vertices for points $x \in \Cset'[X]$. 
\begin{enumerate}[label=(A\arabic*)]
	\setcounter{enumi}{4}
	\item 
	\label{asmp: C'_bounded_density}
	\emph{Bounded density in other clusters:} Letting $\sigma,\lambda_{\sigma}$ be 
	as in \ref{asmp: bounded_density}, for each $\Cset' \in \Cbb_f(\lambda)$ and
	for all $x \in \Csig'$, $\lambda_{\sigma} \leq f(x)$. 
\end{enumerate}

Next we give our main result on consistent cluster recovery by PPR.

\begin{theorem}
	\label{thm: consistent_recovery_of_density_clusters}
	Fix $\lambda > 0$, let $\Cset \in \Cbb_f(\lambda)$ be a
	$\kappa$-well-conditioned density cluster for some $\sigma > 0$, and
	additionally assume $f$ satisfies \ref{asmp: C'_bounded_density}. If Algorithm
	\ref{alg: ppr} is well-initialized with respect to $\Cset$ and if 
	%, there exists a universal constant $c >  0$ and constants in sample size $c_1(\Pbb,r)$-$c_6(\Pbb,r)$ such that if  
	\begin{equation}
	\label{eqn: kappa_ub}
	\kappa(\Cset) \leq c \frac{(\lambda_{\sigma} r^d
		\nu_d)^2}{\vol_{\Pbb,r}(\Csig)}, 
	\end{equation}
	then the following statement holds: there exists a set $\Csig[X]^g \subseteq \Csig[X]$ of large volume 
	with $\vol(\Csig[X]^g; G_{n,r}) \geq \frac{1}{2}\vol(\Csig[X];G_{n,r})$ such that if Algorithm~\ref{alg: ppr} is run with any seed node $v \in \Csig[X]^g$,  then for any $n$ large enough so that
	\begin{equation*}
	n \geq c_1(\Pbb,r)\cdot (\log n)^{\max\{\frac{3}{d},1\}}
	\end{equation*}$\widehat{C}$ satisfies~\eqref{eqn: consistent_density_cluster_recovery} with respect to $\Cset$ with probability at least $1 - \frac{c_2(\Pbb,r)}{n}$. Therefore, $\widehat{C}$ is a consistent estimator of $\Cset$ in the sense of Definition \ref{def: consistent_density_cluster_estimation}. 
\end{theorem}

\subsection{Lower bound.}
\label{sec:lower_bound}

To show a lower bound for density clustering using PPR, we exhibit a hard case: that is, a distribution $\Pbb$ for which PPR is unlikely to recover a density cluster. Let $\mathcal{C}_{\sigma}^{(0)}$, $\mathcal{C}_{\sigma}^{(1)}$, and $\mathcal{C}_{\sigma}^{(2)}$ be rectangles in $\Reals^2$, 
\begin{equation*}
\mathcal{C}_{\sigma}^{(0)} = \biggl[-\frac{\sigma}{2}, \frac{\sigma}{2}\biggr] \times \biggl[-\frac{\rho}{2}, \frac{\rho}{2}\biggr], \quad \mathcal{C}_{\sigma}^{(1)} = \mathcal{C}_{\sigma}^{(0)} - \set{(\sigma,0)}, \quad \mathcal{C}_{\sigma}^{(2)} = \mathcal{C}_{\sigma}^{(0)} + \set{(\sigma,0)} \tag{$0 < \sigma < \rho$}
\end{equation*}
and let $\mathbb{P}$ be the mixture distribution over $\mathcal{X} = \mathcal{C}_{\sigma}^{(0)} \cup \mathcal{C}_{\sigma}^{(1)} \cup \mathcal{C}_{\sigma}^{(2)}$ given by
\begin{equation*}
\mathbb{P} = \frac{1 - \epsilon}{2} \Psi_1 + \frac{1 - \epsilon}{2} \Psi_2 + \frac{\epsilon}{2} \Psi_0,
\end{equation*}
where $\Psi_m$ is the uniform distribution over $\Csig^{(m)}$ for $m = 0,1,2$. 
The density function $f$ of $\Pbb$ is simply
\begin{equation}
\label{eqn:lb_density}
f(x) = \frac{1}{\rho\sigma}\left(\frac{1 - \epsilon}{2}\1(x \in \mathcal{C}_{\sigma}^{(1)}) + \frac{1 - \epsilon}{2}\1(x \in \mathcal{C}_{\sigma}^{(2)}) + \frac{\epsilon}{2}\1(x \in \mathcal{C}_{\sigma}^{(0)})  \right)
\end{equation}
so that for any $\epsilon < \lambda < (1 - \epsilon)/2$, $\mathbb{C}_{f}(\lambda) = \{\mathcal{C}_{\sigma}^{(1)}, \mathcal{C}_{\sigma}^{(2)} \}$.
As the following theorem demonstrates, even when Algorithm~\ref{alg: ppr} is reasonably initialized, if the density cluster $\mathcal{C}_{\sigma}^{(1)}$ is sufficiently geometrically ill-conditioned the cluster estimator $\widehat{C}$ will fail to recover $\mathcal{C}_{\sigma}^{(1)}$. Let
\begin{equation}
\label{eqn:lower_set}
\mathcal{L} = \set{(x_1,x_2) \in \mathcal{X}: x_2 < 0}.
\end{equation}
\begin{theorem}
	\label{thm:ppr_lb}
	Suppose $r < \frac{1}{40}\rho \wedge \frac{1}{4}\sigma$,  $\alpha = 65 \Phi_{\Pbb}(\mathcal{L})$, and $(L,U) = (0,1)$ are inputs to Algorithm~\ref{alg: ppr}. Then, for any
	\begin{equation}
	\label{eqn:lb_sample_size}
	n \geq \max\set{\frac{64}{\epsilon^2 \rho \sigma \pi r^2}, \frac{8}{\epsilon}}
	\end{equation}
	the following statement holds: there exists a set $C^{g} \subset X$ with $\vol_{n,r}(C^g \cap \mathcal{C}_{\sigma}^{(1)}[X]) \geq \frac{1}{10}\vol_{n,r}(\mathcal{C}_{\sigma}^{(1)}[X])$ such that for any seed node $v \in C^g$, the estimator $\widehat{C}$ computed by Algorithm~\ref{alg: ppr} has symmetric set difference with $\Csig^{(1)}[X]$ of volume at least
	\begin{equation}
	\label{eqn:ppr_lb}
	\frac{\sigma \rho}{r^2}\frac{\vol_{n,r}(\widehat{C} \vartriangle \mathcal{C}_{\sigma}^{(1)}[X])}{n^2} \geq \frac{1}{4} -  c \frac{\sqrt{\frac{\sigma}{\rho}}}{\epsilon^2} \sqrt{ \log\left(\frac{\rho \sigma}{\epsilon^2 r^2}\right)\frac{\sigma}{r}}
	\end{equation}
	with probability at least $1 - c_1(\Pbb,r) n \exp\set{-c_2(\Pbb,r)n}$.
	Consequently, if
	\begin{equation*} 
	\epsilon^2 > \frac{c}{8} \sqrt{\frac{\sigma}{\rho}} \cdot \sqrt{ \log\left(\frac{\rho \sigma}{\epsilon^2 r^2}\right)\frac{\sigma}{r}}
	\end{equation*}
	then with high probability $\frac{\sigma \rho}{r^2}\frac{\vol_{n,r}(\widehat{C} \vartriangle \mathcal{C}_{\sigma}^{(1)}[X])}{n^2}$ is at least $1/8$. 
\end{theorem}

The overall takeaway of Theorems~\ref{thm:volume_ssd} and~\ref{thm:ppr_lb} is summarized by Figure~\ref{fig:fig2}, which examines the behavior of PPR clustering over a density concentrated on two semicircles (the ''two-moons'' example). The takeaway is that even in a prototypical example where spectral clustering is designed to work well, PPR will recover high-density clusters (i.e. the moons) only when these clusters are geometrically compact. This is in contrast to traditional plug-in density cluster estimators, which are far more robust to the geometry of $\Pbb$. Figure~\ref{fig:fig2} shows that our theory applies even at moderate sample sizes (in this case $n = 800$ samples).

\subsection{Analysis Overview}
\label{sec: analysis}

One of the primary technical contributions of our work is showing that the geometric
conditions \ref{asmp: bounded_density}--\ref{asmp: bounded_volume} translate to 
meaningful bounds on the normalized cut and mixing time of $\Csig[X]$ in
$G_{n,r}$.

\subsubsection{Upper Bound on the Normalized Cut} 
We start with a finite sample upper bound on the
normalized cut \eqref{eqn: normalized_cut} of $\Cset_\sigma[X]$. For
simplicity, we write $\Phi_{n,r}(\Csig[X]) := \Phi(\Csig[X]; G_{n,r})$.

\begin{theorem}
	\label{thm: conductance_upper_bound}
	Fix $\lambda > 0$, and assume $\Cset \in \Cbb_f(\lambda)$ satisfies
	Assumptions \ref{asmp: bounded_density}--\ref{asmp: low_noise_density}, 
	\ref{asmp: bounded_volume} for some $\sigma > 0$. Then
	\begin{equation}
	\label{eqn: conductance_additive_error_bound}
	\frac{\Phi_{n,r}(\Csig[X])}{r} \leq c_1 \frac{d}{\sigma}
	\frac{\lambda}{\lambda_{\sigma}} \frac{(\lambda_{\sigma} -
		c_0\frac{r^{\gamma}}{\gamma+1})}{\lambda_{\sigma}}
	\end{equation}
	with probability at least $1 - 3\exp\{-nc_3(\Pbb,r)\}$. 
\end{theorem}
\noindent {\bf Remark: } 
Observe that the diameter $\rho$ is absent from Theorem \ref{thm:
	conductance_upper_bound}, in contrast to the condition number
$\kappa(\Cset)$, which worsens (increases) as $\rho$ increases. This
reflects established wisdom regarding spectral partitioning
algorithms more generally \citep{guattery1995, hein2010}, albeit newly applied
to the density clustering setting. It suggests that if the diameter $\rho$ is
large, PPR may fail to recover $\Csig[X]$ even when $\Cset$ is
sufficiently well-conditioned to ensure $\Csig[X]$ has a small normalized
cut in $G_{n,r}$.

\subsubsection{Upper Bounds on the Mixing Time} 
For $S \subseteq V$, denote by $G[S] = (S, E_S)$ the
subgraph induced by $S$ (where the edges are $E_S = E \cap (S \times S)$). Let
$W_S$ be the (lazy) random walk matrix over $G[S]$, and write  
$$
q_{v}^{(t)}(u) = e_vW_S^t e_u
$$
for the $t$-step transition probability of the lazy random walk over $G[S]$
originating at $v \in V$. Also write \smash{$\pi = (\pi(u))_{u \in S}$}
for the stationary distribution of this random walk.(As
$W_S$ is the transition matrix of a lazy random walk, 
it is well-known that a unique stationary distribution exists and is given by 
\smash{$\pi(u) = \deg(u;G[S])/\vol(S; G[S])$}, where we write $\deg(u;G[S]) = \sum_{w \in S} \1((u,w) \in E_S)$ for the degree of $u$ in $G[S]$.) We define the \emph{mixing time} of $G[S]$ as
\begin{equation}
\label{eqn: mixing_time}
\tau_{\infty}(G[S]) = \min\set{ t: \frac{\pi(u) - q_{v}^{(t)}(u)}
	{\pi(u)} \leq \frac{1}{4}, \; \text{for $u,v \in V$}}. 
\end{equation}
Next, we give an asymptotic (in the number of samples $n$) upper bound on
$\tau_{\infty}(G_{n,r}[\Csig[X]])$.

\begin{theorem}
	\label{thm: mixing_time_upper_bound}
	Fix $\lambda > 0$, and assume that \smash{$\Cset \in \Cbb_f(\lambda)$}
	satisfies Assumptions \ref{asmp: bounded_density} and \ref{asmp: embedding}
	for some $\sigma > 0$. Additionally assume that the radius $0 < r < \sigma/2\sqrt{d}$. Then the following statement holds: for any
	\begin{equation*}
	n \geq c_1(\Pbb,r)\cdot(\log n)^{\max\{\frac{3}{d},1\}}
	\end{equation*}
	the mixing time $\tau_{\infty}(G_{n,r}[\Csig[X]])$ satisfies
	\begin{equation} 
	\label{eqn: mixing_time_upper_bound}
	\tau_{\infty}(G_{n,r}[\Csig[X]]) \leq c_2 \frac{\Lambda_{\sigma}^4 d^3 \rho^2 L^2}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{ \Lambda_{\sigma}}{\lambda_{\sigma}^2 r}\right) + c_3
	\end{equation}
	with probability at least $1 - \frac{c_4(\Pbb,r)}{n}$.
\end{theorem}
The proof of Theorem \ref{thm: mixing_time_upper_bound} relies heavily on
analogous mixing time bounds developed for the mixing time of a
continuous-space ``ball walk'' over convex sets. To the best of our knowledge, our
result is the first bound on the mixing time of random walks
over neighborhood graphs that is independent of $n$, the number of vertices.

\begin{figure}
	\centering
	% \begin{adjustbox}{minipage=\linewidth}
	%   \begin{subfigure}{.24\linewidth}
	\includegraphics[width=0.24\textwidth,scale = .5]{example2plots/row1_true_density_cluster}
	% \caption{}
	% \end{subfigure}
	% \begin{subfigure}{.24\linewidth}
	\includegraphics[width=0.24\textwidth]{example2plots/row1_ppr_cluster}
	% \caption{}
	% \end{subfigure}
	% \begin{subfigure}{.24\linewidth}
	\includegraphics[width=0.24\textwidth]{example2plots/row1_conductance_cluster}
	% \caption{}
	% \end{subfigure}
	% \begin{subfigure}{.24\linewidth}
	\includegraphics[width=0.24\textwidth]{example2plots/row1_density_cluster}
	% \caption{}
	% \end{subfigure}

	% \begin{subfigure}{.24\linewidth}
	\includegraphics[width=0.24\textwidth]{example2plots/row2_true_density_cluster}
	% \caption{}
	% \end{subfigure}
	% \begin{subfigure}{.24\linewidth}
	\includegraphics[width=0.24\textwidth]{example2plots/row2_ppr_cluster}
	% \caption{}
	% \end{subfigure}
	% \begin{subfigure}{.24\linewidth}
	\includegraphics[width=0.24\textwidth]{example2plots/row2_conductance_cluster}
	% \caption{}
	% \end{subfigure}
	% \begin{subfigure}{.24\linewidth}
	\includegraphics[width=0.24\textwidth]{example2plots/row2_density_cluster}
	% \caption{}
	% \end{subfigure}

	% \begin{subfigure}{.24\linewidth}
	\includegraphics[width=0.24\textwidth]{example2plots/row3_true_density_cluster}
	% \caption{}
	% \end{subfigure}
	% \begin{subfigure}{.24\linewidth}
	\includegraphics[width=0.24\textwidth]{example2plots/row3_ppr_cluster}
	% \caption{}
	% \end{subfigure}
	% \begin{subfigure}{.24\linewidth}
	\includegraphics[width=0.24\textwidth]{example2plots/row3_conductance_cluster}
	% \caption{}
	% \end{subfigure}
	% \begin{subfigure}{.24\linewidth}
	\includegraphics[width=0.24\textwidth]{example2plots/row3_density_cluster}
	% \caption{}
	% \end{subfigure}
	\caption{\it\small True density (column 1), PPR (column 2), normalized
		cut (column 3) and estimated density (column 4) clusters for 3 different 
		simulated data sets. Seed node for PPR denoted by a black cross.} 
	\label{fig:fig2}
	%\end{adjustbox}
\end{figure}

\section{Hypothesis testing with neighborhood graphs.}

We define test statistics for both the regression and density testing setups. In the former, we show in some preliminary work that our proposed test statistic is minimax optimal over the Sobolev ball $W^{1,2}(\mathcal{X};1)$. In the latter, we empirically demonstrate that our proposed approach has merit. We conclude this section by discussing some practical and computational considerations in this problem. Throughout, we detail the work we intend to complete for a final thesis.

\subsection{Regression testing using graph spectral projections.}
\label{subsec:spectral_regression_testing}

Recalling the regression model~\eqref{eqn:regression}, let $VSV^T$ be the spectral decomposition of the Laplacian matrix $L$ of the neighborhood graph $G_{n,r}$. Inspired by \eqref{eqn:harmonic}, we introduce the following truncated-series test statistic:
\begin{equation}
\label{eqn:graph_spectral_projections}
T_{\mathrm{spec}} := \frac{1}{n} \sum_{k = 0}^{\kappa} \left(\sum_{i = 1} v_i y_i\right)^2
\end{equation}
where the only difference between $T_{\mathrm{harm}}$ and $T_{\mathrm{spec}}$ is that the eigenvectors $\set{v_i}$ are now playing the role of the harmonic functions $\set{\phi_{\ell}}$. In Theorem~\ref{thm:sobolev_testing_rate} we show that under some mild regularity conditions on $\Pbb$, the test $\phi_{\textrm{spec}} := \1\{T_{\mathrm{spec}} \geq \tau\}$ is, up to log factors, a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};1)$. To conveniently state our results we introduce the notation
\begin{equation*}
h(\beta,d) = (1/d + \beta)(d/2 + s)(2d/(4s + d))
\end{equation*}

\begin{theorem}
	\label{thm:sobolev_testing_rate}
	Let $b \geq 1,\beta > 0$ be fixed constants and let $d < 4$. Suppose that $\Pbb$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p$ bounded above and below by constants, i.e
	\begin{equation*}
	0 < p_{\textrm{min}} < p(x) < p_{\textrm{max}} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
	\end{equation*}
	Then there exists a constant $c_1(d,L,p_{\max},b,\beta)$ which is independent of the sample size $n$ such that the following statement holds: Construct the test statistic $T_{\mathrm{spec}}$ with parameter choices $r \asymp n^{-1/d}$ and $\kappa \asymp n^{2d/(4+d)}$, and additionally let the threshold $\tau \asymp \frac{\kappa}{n} + b\sqrt{\frac{\kappa}{n^2}}$. Then, for every $\epsilon$ satisfying
	\begin{equation}
	\label{eqn:sobolev_testing_rate}
	\epsilon^2 \geq c_1(d,L,p_{\max},b,\beta) \cdot b \cdot n^{-4/(4 + d)} (\log n)^{h(\beta,d)}
	\end{equation}
	the worst-case risk is lower bounded
	\begin{equation}
	\label{eqn:sobolev_testing_rate_1}
	\mathcal{R}(\phi_{\mathrm{spec}}; \mathcal{W}^{1,2}(\mathcal{X};L)) \leq \left(\frac{2}{b^2} + \frac{2}{b\sqrt{\kappa}}\right) + o(1).
	\end{equation}
	Therefore, $\epsilon_n^2(\phi_{\mathrm{spec}}, \mathcal{W}^{1,2}(\mathcal{X};L)) \asymp  n^{-4/(4 + d)} (\log n)^{h(\beta,d)}$. 
\end{theorem}

A comparison of Theorem~\ref{thm:sobolev_testing_rate} with the rates derived by~\citet{ingster09} for~\eqref{eqn:harmonic} demonstrates that the additional error incurred by using estimated eigenfunctions of the Laplacian, rather than directly using the Fourier series, results in the critical radius widening by only at most a factor of $(\log n)^{h(\beta,d)/2}$ from the minimax critical radius over $W^{1,2}(\mathcal{X};1)$. This means our test based on $T_{\textrm{spec}}$ is almost \textit{as good as}, in a minimax sense, the best possible test over $W^{1,2}(\mathcal{X};1)$. As mentioned previously, the problem when $d \geq 4$ requires different assumptions on $f$ and is characterized by a different minimax rate.

\subsubsection{Weighted Sobolev norms.}

However, we would like to argue that in certain cases, $T_{\textrm{spec}}$ is in fact \textit{better} than competitors such as $T_{\textrm{harm}}$. While we do not yet have theoretical results along these lines, intuitively $T_{\textrm{spec}}$ should be particularly sensitive to functions $f$ which have smooth deviations from $f_0 = 0$ in high-density regions, but are potentially much wigglier in low-density regions. We therefore intend to (i) alter our function class $\mathcal{H}$ to be defined as the unit ball in a weighted Sobolev norm; precisely
\begin{equation*}
\mathcal{H}_{\Pbb} = \set{f \in L^2(\mathcal{X}): \sum_{\abs{\alpha} \leq s}\int\ \int_{\mathcal{X}} \abs{D^{\alpha}f(x)}^2 p^2(x)\,dx}
\end{equation*}
(ii) analyze the performance of $T_{\mathrm{spec}}$ over $\mathcal{H}_{\Pbb}$, and (iii) provide example distributions $\Pbb$ where $T_{\mathrm{spec}}$ outperforms reasonable competitors over $\mathcal{H}_{\Pbb}$ (See Figure~\ref{fig:fig3} for some empirical examples in the two-sample density testing case).

\subsubsection{Higher-order derivative classes.}
Theorem~\ref{thm:sobolev_testing_rate} covers only the $\{1,2\}$-Sobolev ball. In order to extend to higher-order Sobolev classes (i.e. the Sobolev $\{s,2\}$-balls where $s \geq 2$), we will likely need to consider kernels $K_r(u,v) = K\left(\frac{\norm{u - v}}{r}\right)$ for which
\begin{equation}
\label{eqn:higher_order_kernel}
\int x^i K(x) \,dx = 0, \quad \textrm{i = 1,\ldots,s}
\end{equation}
as is typical in nonparametric smoothing problems \cite{tsybakov08}. Choosing kernels which satisfy~\eqref{eqn:higher_order_kernel} as opposed to $K_r(x,y) = \1(\norm{x - y} \leq r)$ allows for tighter control on the quadratic functional
\begin{equation*}
\int \int (f(x) - f(y))^2 K_r(x,y) \,d\Pbb(x) \,d\Pbb(y);
\end{equation*}
this control is critical to obtain minimax rates, and otherwise the analysis when $s = 1$ versus $s > 1$ is not substantially different. We therefore intend to show that with an appropriate choice of kernel $K$, along with tuning parameters $\kappa$ and $r$, the test statistics $T_{\mathrm{spec}}$ and $T_{\mathrm{spec}}^{(2)}$ (the latter of which we define in the following section) result in minimax optimal tests over all Sobolev $\{s,2\}$-balls, in the regression and density testing problems respectively.

\subsection{Density testing using graph spectral projections.}
\label{subsec:spectral_density_testing}

It is straightforward to adapt the test statistic $T_{\mathrm{spec}}$ to the two-sample testing problem. Concatenate the samples in $X = (z_1,\ldots,z_N,y_1,\ldots,y_M)$, and let $a = (\underbrace{N^{-1},\ldots,{N^{-1}}}_{\textrm{length } N},\underbrace{-M^{-1},\ldots,-M^{-1}}_{\textrm{length } M})$ encode whether a sample in $X$ comes from $Z$ or $Y$. Letting $n = N + M$, we define the variant $T_{\mathrm{spec}}^{(2)}$ of $T_{\mathrm{spec}}$ to be
\begin{equation}
\label{eqn:graph_spectral_projections_2}
T_{\mathrm{spec}}^{(2)} := \frac{1}{n} \sum_{k = 0}^{\kappa} \left(\sum_{i = 1}^{n} v_i a_i\right)^2
\end{equation}
where as before $v_1,\ldots,v_n$ are eigenvectors of the Laplacian $L = L(X,r)$.

While we do not yet have theoretical results characterizing the behavior of $T_{\mathrm{spec}}^{(2)}$, Figure~\ref{fig:fig3} shows some instances in which $T_{\mathrm{spec}}^{(2)}$ outperforms a range of common nonparametric tests. We see that the examples support our intuition; $T_{\textrm{spec}}^{(2)}$ is especially strong when the difference in densities $p - q$ oscillates at a higher frequency in regions where the mixture density $\frac{1}{2}(p + q)$ is small.

\begin{figure}
	\label{fig:fig3}
	\centering
	\includegraphics[width=0.32\textwidth]{plots/sin_density}
	\includegraphics[width=0.32\textwidth]{plots/sin_ROC}
	\includegraphics[width=0.32\textwidth]{plots/sin_theta}
	\includegraphics[width=0.32\textwidth]{plots/stepfunction_sin_density}
	\includegraphics[width=0.32\textwidth]{plots/stepfunction_sin_ROC}
	\includegraphics[width=0.32\textwidth]{plots/stepfunction_sin_theta}
	\includegraphics[width=0.32\textwidth]{plots/stepfunction_sin2_density}
	\includegraphics[width=0.32\textwidth]{plots/stepfunction_sin2_ROC}
	\includegraphics[width=0.32\textwidth]{plots/stepfunction_sin2_theta}
	\includegraphics[width=0.32\textwidth]{plots/stepfunction_sin3_density}
	\includegraphics[width=0.32\textwidth]{plots/stepfunction_sin3_ROC}
	\includegraphics[width=0.32\textwidth]{plots/stepfunction_sin3_theta}
\end{figure}

We intend to (i) derive upper bounds on the worst-case risk of the test $\phi_{\mathrm{spec}}^{(2)} = \1(T_{\mathrm{spec}}^{(2)} \leq \tau)$ over the class $C_d^{s}(\mathcal{X};L)$, and prove that it is minimax optimal, (ii) investigate the minimax rate for two-sample density testing over the Sobolev class $\mathcal{W}_d^{s,2}(\mathcal{X};L)$, and (iii) prove an analogue to Theorem~\ref{thm:sobolev_testing_rate} establish that $\phi_{\mathrm{spec}}^{(2)}$ remains minimax optimal over the larger classes $\mathcal{W}_d^{s,2}(\mathcal{X};L)$.

\subsubsection{Density testing for irregular alternatives.}
As mentioned in Section~\ref{subsubsec:density_testing}, the regression and density testing models over e.g. Holder spaces are formally asymptotically equivalent; however, these asymptotic equivalence statements hold only when $2s > d$. To illuminate some differences between the density and regression testing problems when the regression function $f$ or difference in densities $p - q$ is considered to be irregular (meaning $2s < d$), it is helpful to consider the following generative model: suppose we observe pairs $(x_i,\ell_i)$ for $i = 1,\ldots, (n + m)$, where as usual $x_i \sim \Pbb$ are independent, and additionally $\ell_i$ are conditionally independent given $x_i$ with distribution
\begin{equation*}
\ell_i = 
\begin{cases*}
1, & ~~  \textrm{with probability $\frac{p(x)}{p(x) + q(x)}$} \\
-1, & ~~ \textrm{with probability $\frac{q(x)}{p(x) + q(x)}$.} 
\end{cases*}
\end{equation*}
We can relate this to the density testing problem by applying Bayes rule, whence we derive that the conditional distribution of $X_i|\ell_i = 1$ is $\mathbb{P}$, and similarly the conditional distribution of $X_i|\ell_i = -1$ is $\mathbb{Q}$. On the other hand, since $\ell_i = f(x_i) + w_i$ where $f = \frac{p - q}{p + q}$ and $w_i$ is a mean-zero error term, this resembles in some respects the regression testing setup. However, $w_i$ is now heteroskedastic with variance depending on the unknown densities $p$ and $q$. The fact that the variance of $w_i$ is unknown is crucial, and as previously mentioned it implies that there will no longer be a elbow in rates at $d = 4$. Indeed, as previously stated the minimax critical radius for the two-sample density testing problem over $C_d^1(\mathcal{X};L)$ is known to be $\epsilon_n^{\star}(C_d^1(\mathcal{X};L)) \asymp n^{-2/(4 + d)}$ for all $d$, including $d \geq 4$.

When $d \geq 4$, choosing the hyperparameters for $T_{\mathrm{spec}}^{(2)}$ becomes more subtle. Statistics such as $T_{\mathrm{harm}}$ -- which take empirical evaluations of an orthogonal basis in $L^2(\mathcal{X})$--continue to be minimax optimal over $C_d^{1}(\mathcal{X};L)$ when $\kappa \asymp n^{2d/(4+d)}$. However, since this choice implies $\kappa > n$ the statistic $T_{\mathrm{spec}}^{(2)}$ -- which takes an orthogonal basis in $L^2(\mathbb{P}_n)$ -- reduces to simply $T_{\mathrm{spec}}^{(2)} = 1$ regardless of $\Pbb_n$ and $\Qbb_n$. Intuition from chi-squared tests--where the optimal choice of bin-width is on the order of $n^{-2/(4+d)}$, less than $n^{-1/d}$ whenever $d > 4$--suggests that we should modify our test statistic by choosing $r$ to be shrinking at a rate faster than $n^{-\frac{1}{d}}$. Of course at this connectivity scale with high probability the graph $G_{n,r}$ is disconnected \cite{penrose99}, and our analysis will necessarily be quite different than before. We intend to pursue such a modification and the resulting analysis.

\subsection{IPMs on neighborhood graphs.}
\label{sec:neighborhood_graph_IPMs}

We define a pair of graph-based test statistics on the neighborhood graph $G_{n,r}$, which we cast as IPMs over different ``function'' classes $\Theta \subset \Reals^n$.

\subsubsection{A graph-based Sobolev IPM.}

We term our first test statistic the \emph{graph Sobolev} IPM, since the function class $\Theta$ over which we take the supremum will resemble a continuous Sobolev space. Formally we let
\begin{equation}
\label{eqn:sobolev_IPM}
T_{\textrm{sob}} := \sup_{\theta \in \Reals^n: \norm{B^{(s)}\theta}_2 + \lambda\norm{\theta}_2 \leq 1} \abs{a^T \theta} = \sqrt{a^T (L^s + \lambda I)^{-1}a}, ~~ \lambda > 0, 
\end{equation}
where we recall the notation $a = \left(N^{-1},\ldots,N^{-1},-M^{-1},\ldots,-M^{-1}\right)$. (We could easily adapt this statistic to the regression testing case by replacing $a$ with $y$, although we would lose the analogy to an IPM).

\textbf{Remark:}~~Two aspects of the statistic $T_{\textrm{sob}}$ are worthy of comment. The first regards the role played by the parameter $\lambda$, which is at first glance somewhat mysterious. It may seem more natural to consider the statistic
\begin{equation*}
\sup_{\theta \in \Reals^n: \norm{B\theta}_2 \leq 1} \abs{a^T \theta}, 
\end{equation*}
so that the function class is defined only with respect to the semi-norm $\norm{B\theta}_2$. However, it can be shown that this function class is too rich, and our test statistic will overfit to noise; hence we restrict it in the manner of~\eqref{eqn:sobolev_IPM}, a modification similar to that of \citet{arbel18} who observe that it improves empirical performance. Second, we note that in general, an analogous IPM in the continuum Sobolev space is a poor candidate for testing. Specifically, we have that when $2s < d$,
\begin{equation}
\label{eqn:continuum_Sobolev_IPM}
\sup_{f \in L^2(\mathcal{X}): \norm{f}_{W^{s,2}(\mathcal{X})} \leq L} \abs{\mathbb{P}_nf - \mathbb{Q}_nf} = \infty
\end{equation}
regardless of $\mathbb{P}_n$ and $\mathbb{Q}_n$, achieved by taking spike functions at data. (This phenomenon has been observed in the context of semi-supervised learning in e.g. \citet{nadler09,zhou11,slepcev17,alaoui16}.) However, by replacing the continuous Sobolev norm in \eqref{eqn:continuum_Sobolev_IPM} with a discrete Sobolev norm in \eqref{eqn:sobolev_IPM}, we have obviated this concern.

Based on the success of Sobolev IPMs in real-world problems \citet{mroueh17,arbel18} and the favorable properties of analogous estimators when performing regression on a grid \citet{sadhanala16}, we believe testing with $T_{\textrm{sob}}$ is a reasonable idea. We intend to investigate the minimax properties of such a test with respect to the function classes $C^s(\mathcal{X})$ and $W^{s,2}(\mathcal{X})$.

\subsubsection{A graph-based Total Variation (TV) IPM.}
Our second IPM test statistic will resemble the first, except we replace the graph Sobolev norm by a graph total variation norm. We hence call this statistic the \emph{graph TV} IPM, defined as
\begin{equation*}
T_{\mathrm{TV}} := \sup_{\theta \in \Reals^n: \norm{B\theta}_1 + \lambda\norm{\theta}_2 \leq 1} \abs{a^T \theta}.
\end{equation*}
To motivate the graph TV IPM, we give some background on testing in TV spaces.

\paragraph{Testing in TV spaces.}

To one extent or another, functions in the Sobolev and Holder balls exhibit homogeneous smoothness. In reality we may not wish to restrict our attention to such functions. This is in part the motivation behind considering functions of bounded variation. Formally, for a function $f \in L^1(\mathcal{X})$ the \emph{total variation} semi-norm of $f$ is \citet{evans15}
\begin{equation*}
TV(f;\mathcal{X}) := \sup \left\{ \int_U f \, \dive \, \psi \,dx : \psi \in C_c^1(\mathcal{X}; \Reals^d), \abs{\psi} \leq 1 \right\};
\end{equation*}
and we write $BV_d(\mathcal{X})$ for the subset of functions $f \in L^1(\mathcal{X})$ which have bounded norm
\begin{equation*}
\norm{f}_{BV_d(\mathcal{X})} := \norm{f}_{\infty} + TV(f;\mathcal{X}).
\end{equation*}

In the one-dimensional case, some work has been done on nonparametric testing over the Besov spaces $B_{p,q}^s(\mathcal{X})$ (for a formal definition see Section 4.3 of \cite{gine16}), which we can relate to bounded variation spaces via the embeddings
\begin{equation}
\label{eqn:besov_embeddings}
B_{11}^1(\mathcal{X}) \subseteq BV_1(\mathcal{X}) \subseteq B_{1\infty}^1(\mathcal{X}).
\end{equation}

\citet{lepski1999} study the regression testing problem, with data observed according to the white noise model as in \eqref{eqn:gaussian_white_noise}. They derive among other things that for the Besov balls $B_{1,q}^s(\mathcal{X}; L)$, the minimax critical radius $\epsilon_n^{\star}(B_{1,q}^s(\mathcal{X}; L)) \asymp n^{-2s'/(4s' + 1)}$, where $s' = s - \frac{1}{4}$ and the result holds for any $s > 1, q \geq 1$. \citet{ingster00} improve on this by showing the same result holds whenever $s > 1/2$. This along with~\eqref{eqn:besov_embeddings} implies that $\epsilon_n^{\star}(BV_1(\mathcal{X})) \asymp n^{-\frac{3}{8}}$.

When $d \geq 2$ minimax testing rates over the Besov and bounded variation spaces are still unknown. However, study of the equivalent estimation problem suggests the answers may be substantially different from the 1d problem. \citet{sadhanala16} consider a discretized version of this problem -- where observations are made on a discrete grid and the regression function is assumed to lie within a discrete bounded variation class -- and show that the minimax estimation rate in empirical norm is on the order of $n^{-1/d}$ for $d \geq 2$. This is slower than minimax estimation rate over the Sobolev space $\mathcal{W}_d^{1,2}$, the largest Sobolev space inscribed within $BV_d$. \citet{delyon96,kerkyacharian08,lepski2015,ruiz18} all consider the regression white noise model \eqref{eqn:gaussian_white_noise} and show, in progressive degrees of generality, that the minimax rates over the Besov spaces $B_{p,q}^s$ exhibit a phase transition at the boundary $2s = d$. When $2s > d$, the minimax estimation rate (with loss measured in squared $\mathcal{L}_2$ norm) is the standard $n^{-2s/(2s + d)}$; on the other hand when $2s < d$, the minimax rate is $n^{-s/d}$ which is much slower. When $s = 1$, this phase transition occurs at $d = 2$, and so in a minimax sense the univariate and multivariate estimation problems over $BV_d$ are revealed to be quite different. \citet{sadhanala17} show that this same phenomenon carries over to the grid, and the discrete TV classes defined upon it.

We intend to investigate whether such a phase transition (from $d = 1$ to $d \geq 2$) exists in the testing setting as well, by deriving upper and lower bounds for the critical radius $\epsilon_n(BV_d(\mathcal{X};L))$, where $BV_d(\mathcal{X}; L) = \set{f \in BV_d(\mathcal{X}): \norm{f}_{BV_d(\mathcal{X})} \leq L}$. Our upper bound will be based on the test $\phi_{TV} := \1\{T_{\mathrm{TV}} \geq \tau \}$.

\subsubsection{Computational considerations.}

Although our primary focus has been on the statistical properties of various tests we have defined, there are also some computational considerations which merit our attention.

When $n$ is large and the graph $G_{n,r}$ is relatively dense, exactly computing the eigenvectors of $L$ (as in~\eqref{eqn:graph_spectral_projections}) or the matrix inverse $(L + \lambda I)^{-1}$ (as in ~\eqref{eqn:sobolev_IPM}) may be computationally expensive. Recently, there has been remarkable progress made (e.g. \citet{spielman2011,batson13,spielman2014,cohen14}) in efficiently computing $\epsilon$-\emph{spectral sparsifiers} of $G$: that is, graphs $\widetilde{G}$ with Laplacian matrices $\widetilde{L}$ which satisfy:
\begin{equation}
\label{eqn:spectral_similarity}
(1 - \epsilon)x^T \widetilde{L} x \leq x^T L x \leq (1 + \epsilon) x^T \widetilde{L} x, \quad \textrm{for some $0 \leq \epsilon < 1$, and all $x \in \Reals^n$.}
\end{equation}
and have many fewer edges than $G$ (hence the term ``sparsifier''.) Since $\widetilde{L}$ is a sparse matrix, we can compute its eigenvectors $\set{\widetilde{v}}$ or the inverse $(\widetilde{L} + \lambda I)^{-1}$  efficiently. Additionally \eqref{eqn:spectral_similarity} should imply that the resulting test statistics behave similarly to those defined with respect to $L$. As yet limited work \citet{vonluxburg2014,sadhanala16b} has been done to rigorously analyze the performance of spectral sparsifiers in statistical learning problems, despite their obvious promise. We plan to address this gap by providing a theoretical characterization of the performance of graph statistics on sparsified graphs in the regression and density testing problems.

% We also mention in passing that another potential computational difficulty stems from calibrating the threshold $\tau$. While the choice of $\tau$ in the statement of Theorem~\ref{thm:sobolev_testing_rate} is optimal in a minimax sense, practically speaking depending on the regression function $f$ and distribution $\Pbb$ it may be a conservative threshold, with a resulting loss in power. In practice, one can calibrate $\tau$ by permutation, however doing so requires repeatedly shuffling the data $X$ and recomputing a test statistic $T$. It would therefore be helpful to know the distribution of our various proposed statistics $T$ under the null hypothesis, or at least derive asymptotic approximations to these distribution, and we plan to investigate this. However, while in the density testing context there has been some work on deriving asymptotic distributions of IPMs \cite{gretton12,sadhanala19}, typically these limiting distributions depend on the unknown $\mathbb{P}$, and therefore the practical implications of these results is somewhat limited.

\clearpage

\bibliographystyle{plainnat}
\bibliography{proposal_bibliography}

\end{document}

### 2019-11-19 notes/proposal/proposal_document.tex
modified proposal doc based on Siva suggestions

@@
 	In the classical data analysis setting, where we observe i.i.d.
-point-cloud data sampled from an unknown distribution $P$, there has been recent interest in the design and analysis of procedures which involve the geometry of $P$.
-Broadly speaking, these procedures either estimate geometric properties such as high-density regions of $P$, or leverage the geometry of $P$ in some downstream statistical learning task.
-One way to implicitly adapt to the geometry of $P$ is through the use of graph-based spectral methods.
-The underlying intuition is that when one appropriately constructs a neighborhood graph, the resulting graph Laplacian encodes a great deal of information about the geometry of $P$; for example, one can view its top eigenvectors as (discrete versions of) functions which are smooth with respect to $P$.
+point-cloud data sampled from an unknown distribution $\Pbb$, there has been recent interest in the design and analysis of procedures which involve the geometry of $\Pbb$.
+Broadly speaking, these procedures either estimate geometric properties such as high-density regions of $\Pbb$, or leverage the geometry of $\Pbb$ in some downstream statistical learning task.
+One way to implicitly adapt to the geometry of $\Pbb$ is through the use of graph-based spectral methods.
+The underlying intuition is that when one appropriately constructs a neighborhood graph, the resulting graph Laplacian encodes a great deal of information about the geometry of $\Pbb$; for example, one can view its top eigenvectors as (discrete versions of) functions which are smooth with respect to $\Pbb$.
 For this reason, there has been extensive analysis on consistency properties of graph Laplacians, that is, in what ways and with what error they approximate continuum-level operators.
@@
 
-\section{Background.}
-
-\textcolor{red}{TODO:} This is why you are going to read.
-
-\textcolor{red}{TODO:} This is what you are going to read.
+\section{Introduction and background.}
+
+This document details the work I intend to complete for my thesis, and is separated into three sections.
+In the first section, I give background on the convergence properties of neighborhood graphs, as well as the statistical learning problems of clustering and hypothesis testing.
+In the second section, I detail work I have already completed on clustering using the local spectral algorithm PPR over neighborhood graphs.
+In the third section, I propose several statistics used to test hypotheses using the spectra of neighborhood graphs, and make explicit the ways in which I will investigate and characterize their behavior.
+Finally, I give an estimated timeline for completing my proposed work.
 
@@
 Let $X = \{x_1,
-\ldots, x_n\}$ be a sample drawn i.i.d.\ from a distribution $\Pbb$ on $\Rd$,
+\ldots, x_n\}$ be a sample drawn i.i.d. from a distribution $\Pbb$ on $\Rd$,
 with density~$f$.  For a radius $r > 0$, we define $G_{n,r}=(V,E)$ to be the
@@
 
-Spectral graph theory involves the study of the eigenvalues and eigenvectors of any of following Laplacian matrices:
-\begin{equation*}
-\textrm{combinatorial:}~ L = D - A,~~ \textrm{random walk:}~ L_{\textrm{rw}} = I - D^{-1}A,~~ \textrm{normalized:}~ L_{\textrm{n}} = D^{1/2}L_{\textrm{rw}}D^{-1/2}.
-\end{equation*}
-
-and is broadly applicable to fields as diverse as the theory of randomized algorithms and differential geometry \textcolor{red}{Chung}.
-When specifically considering Laplacians of neighborhood graphs such as $G_{n,r}$, one well-developed line of work concerns the asymptotic convergence properties of the various graph Laplacians to associated continuum Laplacian operators.
-
-\begin{itemize}
-	\item  \textcolor{red}{Define the Laplacian operator}.
-	\item Introduce the study of Laplacians on point-cloud data, manifold learning, and give references as to their background.
-Discuss the study of the convergence of spectra \textcolor{red}{Belkin, Coifman, Garcia-Trillos and Slepcev 1, Garcia-Trillos and Slepcev Spec, von Luxburg, Hein + von Luxburg, Lafon,Koltchinskii and Gine, Gine and Koltchinskii 1}, pointwise convergence of operators \textcolor{red}{Belkin thesis, Lafon thesis, Hein + Audibert, Singer, Belkin + Niyogi08}, and convergence of smoothness functionals \textcolor{red}{Bousquet + Chapelle, Hein07}.
-	\item Discuss the study of other functionals, such as total variation and normalized cut, on graphs \textcolor{red}{(Garcia-Trillos and Slepcev TV, Arias-Castro, Maier + von Luxburg)}
-	\item 
-\end{itemize}
-
-Why should I care about the asymptotic convergence of graph Laplacians?
-\begin{itemize}
-	\item The convergence of graph Laplacians, and their spectra, to underlying continuum operators suggests they may be used to perform a variety of desirable tasks in statistical learning. 
-	\item Dimension reduction: under the hypothesis that the data lie on, or are concentrated near, a manifold $\mathcal{M} \subset \Reals^d$ of dimension $m < d$, the eigenvectors $v_1,\ldots,v_m$, can be used to embed $x_i \mapsto (v_{i,1},\ldots v_{i,m})$, reducing the dimension from $d$ to $m$ without losing the geometric structure of $\Pbb$.
-This algorithm is known as Laplacian Eigenmaps (Belkin).
-	\item \textcolor{red}{Clustering: under the hypothesis that the distribution $\Pbb$ is supported on disjoint connected components $\mathcal{C}_1,\ldots,\mathcal{C}_k$, the span of the smallest $k$ eigenvectors of the Laplacian operator $L_n$ will approximately equal the span of the functions $\phi_j(x) = \1\set{x \in \mathcal{C}_j}, j = 1,\ldots,k$.} \textcolor{red}{(Belkin + Arias-Castro)}
-	\item Semi-supervised and supervised learning: Discuss the study of Laplacians in supervised learning \textcolor{red}{(Bousquet + Chapelle, Lee + Izbicki, GarciaTrillos + Murray, Padilla)} Discuss the study of Laplacians in semi-supervised learning \textcolor{red}{(Belkin + Niyogi, Zhou + Belkin, Slepcev and Thorpe)}.
-\end{itemize}
-
-In this thesis proposal, we pick up a few of these threads, showing how spectral methods can be used to perform the statistical tasks of density clustering and nonparametric hypothesis testing.
+Spectral graph theory involves the study of the eigenvalues and eigenvectors of the graph Laplacian $L := D - A$\footnote{Also the normalized Laplacian matrices $L_{\textrm{rw}} = I - D^{-1}A$ and $L_{\textrm{n}} = D^{1/2}L_{\textrm{rw}}D^{-1/2}$.} and is broadly applicable to fields as diverse as the theory of randomized algorithms and differential geometry \citet{chung97}.
+
+When specifically considering Laplacians of neighborhood graphs such as $G_{n,r}$, one well-developed line of work concerns the asymptotic convergence properties of the various graph Laplacians to associated continuum Laplace operators.
+Let $\mathcal{X} \subseteq \Reals^d$ be an open, bounded, and connected set with Lipschitz boundary.
+The Laplace operator $\Delta: C^2(\mathcal{X}) \to L^2(\mathcal{X})$ is defined as
+\begin{equation*}
+\Delta g = - \dive(\nabla g) =: \sum_{i = 1}^{d} \frac{\partial^2 g}{\partial x_i^2}.
+\end{equation*}
+There are several ways in which the matrix $L$ can be said to be converging to the operator $\Delta$.
+The first of these is \emph{functional} convergence.
+Here, we think of $L$ as acting on functions in $\Reals^d$,
+\begin{equation*}
+Lg(x) := \frac{1}{n^2r_n^{d+2}}\sum_{i = 1}^{n}(g(x) - g(x_i))A_{ij}
+\end{equation*}
+and ask whether $Lg(x) \to \Delta g(x)$ in probability as $n \to \infty$ and $r_n \to 0$.
+\citet{belkin03,belkin05} consider a slightly different neighborhood graph than $G_{n,r}$ (constructed with a Gaussian rather than uniform kernel) and show that when $\Pbb$ is the uniform distribution over $\mathcal{X}$, such a functional convergence occurs.
+\footnote{For this and many of the results summarized in this section, $\mathcal{X}$ is permitted to be a low-dimensional manifold, and $\Delta$ is then taken to be the Laplace-Beltrami operator.} \citet{singer06,gine06} extend these results by proving faster convergence rates and a uniform CLT, respectively.
+When $\Pbb$ is non-uniform the story is slightly different.
+Now the relevant continuum operator is a weighted analogue of the Laplace operator,
+\begin{equation*}
+\Delta_{\Pbb} g := - \frac{1}{f} \dive(f^2 \nabla g);
+\end{equation*}
+\citet{lafon04,hein05} prove the functional convergence $Lg(x) \to \Delta_{\Pbb} g(x)$ in probability as $n \to \infty$ as $n \to \infty$ and $r_n \to 0$.
+
+Another mode of convergence is \emph{smoothness functional} convergence. In this case, we denote $\theta = (g(x_1),\ldots,g(x_n)) \in \Reals^n$, consider the smoothness functional
+\begin{equation*}
+S(g;\Pbb) := \int_{\mathcal{X}} \norm{\nabla g}_2^2~ f^2(x) \,dx
+\end{equation*}
+and ask whether $\frac{1}{n^2r_n^{d+2}}\theta^T L \theta \to S(g;\Pbb)$ in probability as $n \to \infty$ and $r_n \to 0$.
+\citet{bousquet03} proves that the smoothness functional converges pointwise--meaning for a fixed $g \in C^2(\mathcal{X})$--and \citet{hein06} extends this to show uniform convergence over a suitably chosen Holder ball.
+
+The final mode of convergence we mention is \emph{spectral} convergence.
+$\Delta_{\Pbb}$ is a positive semidefinite self-adjoint operator, and therefore has a discrete spectrum.
+Label the $i$th smallest eigenvalue of $\Delta_{\Pbb}$ as $\lambda_i$, with associated eigenfunction $u_i$; similarly, label the $i$th smallest eigenvalue of $L$ as $s_i$ with associated eigenvector $v_i$.
+We say $L$ converges spectrally to $\Delta_{\Pbb}$ if for any fixed $k \in \mathbb{N}$,
+\begin{equation*}
+\lim_{n \to \infty} s_{k} = \lambda_k, \quad \textrm{and} \quad \lim_{n \to \infty} \frac{1}{n}\sum_{i = 1}^{n} (u_k(x_i) - v_{k,i})^2 = 0
+\end{equation*}
+where the limits are taken in probability.
+\citet{vonluxburg2008} consider the case of a fixed $r_n = r$, and prove spectral convergence of $L$ to an appropriately defined integral operator, \citet{belkin07} prove that when $\Pbb$ is the uniform distribution over $\mathcal{X}$, $L$ converges spectrally to $\Delta_{\Pbb} = \Delta$ for some $r_n \to 0$ as $n \to \infty$, and \citet{garciatrillos18} extend this result to hold for general $\Pbb$ and obtain rates at which $r_n$ may converge to $0$.
+
+The convergence of graph Laplacians to underlying continuum differential operators suggests they may be used to perform a variety of desirable tasks in statistical learning.
+In this thesis proposal, we study a few of these tasks in detail, and show how spectral methods can be used to perform the statistical tasks of density clustering and nonparametric hypothesis testing.
 Before doing so, we provide some background on each of these areas.
+
+%\begin{itemize}
+	%\item Dimension reduction: under the hypothesis that the data lie on, or are concentrated near, a manifold $\mathcal{M} \subset \Reals^d$ of dimension $m < d$, the eigenvectors $v_1,\ldots,v_m$, can be used to embed $x_i \mapsto (v_{i,1},\ldots v_{i,m})$, reducing the dimension from $d$ to $m$ without losing the geometric structure of $\Pbb$.
+This algorithm is known as Laplacian Eigenmaps \citet{belkin03a}.
+	%\item Clustering: under the hypothesis that the distribution $\Pbb$ is approximately supported on disjoint connected components $\mathcal{C}_1,\ldots,\mathcal{C}_k$, the embedding $x_i \mapsto (v_{i,1},\ldots,v_{i,k})$ will reveal the cluster structure of $\Pbb$.
+Subsequent application of a simple clustering method such as $k$-means clustering will then correctly partition the data $X$ into clusters \cite{ng01,ariascastro11,shi2009}.
+	%\item Semi-supervised and supervised learning: Discuss the study of Laplacians in supervised learning \textcolor{red}{(Bousquet + Chapelle, Lee + Izbicki, GarciaTrillos + Murray, Padilla)} Discuss the study of Laplacians in semi-supervised learning \textcolor{red}{(Belkin + Niyogi, Zhou + Belkin, Slepcev and Thorpe)}.
+%\end{itemize}
 
@@
 \citet{hartigan1981, chaudhuri2010, balakrishnan2013,kpotufe11} consider consistent estimation of the
-cluster tree. None of these works, however, study the ability of spectral algorithms to recover density clusters, which will be our focus.
+cluster tree. None of these works, however, study the ability of spectral algorithms to recover density clusters, which will be our focus in Section~\ref{sec:density_clustering_PPR}.
 
@@
 
-The results in Section~\ref{subsec:neighborhood_graphs_spectra} detailing with spectral convergence of the graph Laplacian operator have straightforward consequences for the consistency of spectral clustering (for instance, see \cite{vonluxburg2008,garciatrillos18}).
+The results in Section~\ref{subsec:neighborhood_graphs_spectra} have straightforward consequences for the consistency of spectral clustering (for instance, see \cite{vonluxburg2008,garciatrillos18}).
 However, the behavior of the spectra of these continuum operators can in general be hard to grasp.
@@
 
-In addition to observing i.i.d samples $x_1,\ldots,x_n$ from a distribution $\mathbb{P}$ with support $\mathcal{X} := [0,1]^d \subset \Reals^d$\footnote{For the purposes of this proposal, whenever we discuss hypothesis testing we will assume $\mathcal{X} = [0,1]^d$, i.e.
+In addition to observing i.i.d samples $x_1,\ldots,x_n$ from a distribution $\mathbb{P}$ with density $p$ supported on $\mathcal{X} := [0,1]^d \subset \Reals^d$\footnote{For the purposes of this proposal, whenever we discuss hypothesis testing we will assume $\mathcal{X} = [0,1]^d$, i.e.
 the unit cube with dimension equal to that of the ambient space $\Reals^d$.}, suppose we observe $Y = \set{y_1,\ldots,y_n}$ according to the following regression model:
@@
 \begin{equation*}
-\epsilon_n^{\star}(\mathcal{H}) = \inf \set{\epsilon:\inf_{\phi} \mathcal{R}(\phi; \mathcal{H}_{\epsilon} ) \leq \alpha}.
+\epsilon_n^{\star}(\mathcal{H}) = \inf \set{\epsilon:\inf_{\phi} \mathcal{R}(\phi; \mathcal{H}_{\epsilon} ) \leq \alpha},
 \end{equation*}
@@
 
-\paragraph{Testing over Sobolev classes.}
+\paragraph{Testing over Sobolev spaces.}
 The first type of function class $\mathcal{H}$ we consider will be balls in Sobolev spaces.
-We say a Borel measurable function $f: \mathcal{X} \to \Reals$ is in the space $\mathcal{L}^p(\mathcal{X})$ for $1 \leq p < \infty$ if $\int_{\mathcal{X}} \abs{f(x)}^p \,dx < \infty$.
-Then, the Sobolev space $W_d^{s,2}(\mathcal{X})$ consists of all functions $f \in \mathcal{L}^1(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$(See \citet{evans10} for a formal definition) belongs to $\mathcal{L}^p(\mathcal{X})$.
+We say a Borel measurable function $f: \mathcal{X} \to \Reals$ is in the space $\mathcal{L}^p(\mathcal{X})$ for $1 \leq p < \infty$ if
+$\norm{f}_{\mathcal{L}^p(\mathcal{X})} := \int_{\mathcal{X}} \abs{f(x)}^p \,dx < \infty$.
+Then, the Sobolev space $W_d^{s,2}(\mathcal{X})$ consists of all functions $f \in \mathcal{L}^2(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ (See \citet{evans10} for a formal definition) belongs to $\mathcal{L}^2(\mathcal{X})$.
 The Sobolev $\{s,2\}$ norm is then
@@
 \end{equation*}
-and for a given $L > 0$, the corresponding ball is $W_d^{s,2}(\Xset; L) = \set{f: \norm{f}_{W^{s,2}(\Xset)} \leq L}$. In this work, we will treat $L$ as a fixed, known constant.
+and for a given $L > 0$, the corresponding ball is $W_d^{s,2}(\Xset; L) = \set{f: \norm{f}_{W^{s,2}(\Xset)} \leq L}$. (In this work, we will treat $L$ as a fixed, known constant.)
 
@@
 The case where we observe data according to the generative model~\eqref{eqn:regression} was treated in \cite{ingster09}, wherein it is shown that the minimax rate remains the same  $\epsilon_n^{\star}(\mathcal{W}_d^{s,2}(\mathcal{X}, L)) \asymp n^{-2s/(4s + d)}$ so long as $4s > d$.
-When $4s < d$, if we replace the assumption $f \in W_d^{s,2}(\mathcal{X},L)$ by the assumption $\left(\int_{\mathcal{X}} f^4(x) \,dx\right)^{1/4}$, then \citet{guerre02} show that a test $\phi$ based on the statistic $\frac{1}{n}\sum_{i = 1}^{n}y_i^2$ will have testing error on the order of $n^{-1/4})$ .
-Since $\mathcal{W}^{s,2}$ does not compactly embed into $\mathcal{L}^4$ when $4s < d$ this result does not apply to Sobolev spaces in this regime; indeed it seems likely that uniformly consistent testing over these spaces is not possible, as the function $f^2$ may no longer have finite variance.
+When $4s < d$, if we replace the assumption $f \in W_d^{s,2}(\mathcal{X},L)$ by the assumption $\left(\int_{\mathcal{X}} f^4(x) \,dx\right)^{1/4}$, then \citet{guerre02} show that a test $\phi$ based on the statistic $\frac{1}{n}\sum_{i = 1}^{n}y_i^2$ will have testing error on the order of $n^{-1/4}$ .
+Since $\mathcal{W}^{s,2}$ does not compactly embed into $\mathcal{L}^4$ when $4s < d$ this result does not apply to Sobolev spaces in this regime; indeed it seems likely that no uniformly consistent test exists over these spaces, as the function $f^2$ may no longer have finite variance.
+
+\paragraph{Testing over Total Variation (TV) spaces.}
+
+To one extent or another, functions in the Sobolev balls exhibit homogeneous smoothness.
+In reality we may not wish to restrict our attention to such functions.
+We therefore consider functions of bounded total variation.
+Formally, for a function $f \in L^1(\mathcal{X})$ the \emph{total variation} semi-norm of $f$ is \citet{evans15}
+\begin{equation*}
+TV(f;\mathcal{X}) := \sup \left\{ \int_{\mathcal{X}} f \, \dive \, \psi \,dx : \psi \in C_c^1(\mathcal{X}; \Reals^d), \abs{\psi} \leq 1 \right\};
+\end{equation*}
+and we write $BV_d(\mathcal{X})$ for the subset of functions $f \in L^1(\mathcal{X})$ which have bounded norm
+\begin{equation*}
+\norm{f}_{BV_d(\mathcal{X})} := \norm{f}_{\infty} + TV(f;\mathcal{X}).
+\end{equation*}
+
+In the one-dimensional case, some work has been done on nonparametric testing over the Besov spaces $B_{p,q}^s(\mathcal{X})$ (for a formal definition see Section 4.3 of \cite{gine16}), which we can relate to the bounded variation space $BV_1(\mathcal{X})$ via the embeddings
+\begin{equation}
+\label{eqn:besov_embeddings}
+B_{11}^1(\mathcal{X}) \subseteq BV_1(\mathcal{X}) \subseteq B_{1\infty}^1(\mathcal{X}).
+\end{equation}
+
+\citet{lepski1999} study the regression testing problem, with data observed according to the white noise model as in \eqref{eqn:gaussian_white_noise}.
+They derive among other things that for the Besov balls $B_{1,q}^s(\mathcal{X}; L)$, the minimax critical radius $\epsilon_n^{\star}(B_{1,q}^s(\mathcal{X}; L)) \asymp n^{-2s'/(4s' + 1)}$, where $s' = s - \frac{1}{4}$ and the result holds for any $s > 1, q \geq 1$.
+\citet{ingster00} improve on this by showing the same result holds whenever $s > 1/2$.
+This along with~\eqref{eqn:besov_embeddings} implies that $\epsilon_n^{\star}(BV_1(\mathcal{X})) \asymp n^{-\frac{3}{8}}$.
+
+When $d \geq 2$ minimax testing rates over the Besov and bounded variation spaces are still unknown.
+However, study of the equivalent estimation problem suggests the answers may be substantially different from the 1d problem.
+\citet{sadhanala16} consider a discretized version of this problem -- where observations are made on a discrete grid and the regression function is assumed to lie within a discrete bounded variation class -- and show that the minimax estimation rate in empirical norm is on the order of $n^{-1/d}$ for $d \geq 2$.
+This is slower than minimax estimation rate over the Sobolev space $\mathcal{W}_d^{1,2}$, the largest Sobolev space inscribed within $BV_d$.
+\citet{delyon96,kerkyacharian08,lepski2015,ruiz18} all consider the regression white noise model \eqref{eqn:gaussian_white_noise} and show, in progressive degrees of generality, that the minimax rates over the Besov spaces $B_{p,q}^s$ exhibit a phase transition at the boundary $2s = d$.
+When $2s > d$, the minimax estimation rate (with loss measured in $\mathcal{L}^2$ norm) is the standard $n^{-s/(2s + d)}$; on the other hand when $2s < d$, the minimax rate is $n^{-s/2d}$ which is much slower.
+When $s = 1$, this phase transition occurs at $d = 2$, and so in a minimax sense the univariate and multivariate estimation problems over $BV_d$ are revealed to be quite different.
+\citet{sadhanala17} show that this same phenomenon carries over to the grid, and the discrete TV classes defined upon it.
 
@@
 (Although we overload notation by having $\mathcal{R}$ and $\epsilon_n$ refer to the worst-case risk and critical radius in both the regression and density testing problems, it will always be clear from context which problem we are referring to.)
-\paragraph{Density testing in Holder spaces.} Unlike in the regression testing context, where $f$ is assumed to belong to a Sobolev space, most of the study of density testing seems to have dealt with the case where $\mathcal{H}$ is Holder space.
+
+\paragraph{Testing over Holder spaces.} Unlike in the regression testing context, where $f$ is assumed to belong to a Sobolev space, most of the study of density testing seems to have dealt with the case where $\mathcal{H}$ is a Holder space.
 For a given $s > 0$, the $s$th Holder norm is given by
@@
 $\kappa \in \mathbb{N}$ a tuning parameter.
-The test statistic $T_{\mathrm{spec}}$ which we propose and analyze in Section~\ref{subsec:spectral_regression_testing} is closely related to $T_{\mathrm{Harm}}$, the crucial difference being that instead of the Fourier basis $\set{\phi_{\ell}}$, which are eigenfunctions of the Laplacian operator over $[0,1]^d$, we use eigenvectors $\set{v_k}$ of the graph Laplacian matrix.
+The test statistic $T_{\mathrm{spec}}$ which we propose and analyze in Section~\ref{subsec:spectral_regression_testing} is closely related to $T_{\mathrm{Harm}}$, the crucial difference being that instead of the Fourier basis $\set{\phi_{\ell}}$, which are eigenfunctions of the Laplacian operator $\Delta$, we use eigenvectors $\set{v_k}$ of the graph Laplacian matrix.
 
@@
 \section{Density clustering with PPR.}
+\label{sec:density_clustering_PPR}
 
@@
 In the latter, we empirically demonstrate that our proposed approach has merit.
-We conclude this section by discussing some practical and computational considerations in this problem.
-Throughout, we detail the work we intend to complete for a final thesis.
+We conclude this section by discussing some practical and computational considerations for this problem.
+Throughout, in subsections entitled ``Proposed Work'' we detail the work we intend to complete for a final thesis.
 
@@
 Recalling the regression model~\eqref{eqn:regression}, let $VSV^T$ be the spectral decomposition of the Laplacian matrix $L$ of the neighborhood graph $G_{n,r}$.
-Inspired by \eqref{eqn:harmonic}, we introduce the following truncated-series test statistic:
+Inspired by \eqref{eqn:harmonic}, we propose the following truncated-series test statistic:
 \begin{equation}
@@
 \end{equation}
-where the only difference between $T_{\mathrm{harm}}$ and $T_{\mathrm{spec}}$ is that the eigenvectors $\set{v_i}$ are now playing the role of the harmonic functions $\set{\phi_{\ell}}$.
+where the main difference between $T_{\mathrm{spec}}$ and $T_{\mathrm{harm}}$ is that the eigenvectors $\set{v_i}$ are now playing the role of the harmonic functions $\set{\phi_{\ell}}$.
 In Theorem~\ref{thm:sobolev_testing_rate} we show that under some mild regularity conditions on $\Pbb$, the test $\phi_{\textrm{spec}} := \1\{T_{\mathrm{spec}} \geq \tau\}$ is, up to log factors, a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};1)$.
@@
 
-A comparison of Theorem~\ref{thm:sobolev_testing_rate} with the rates derived by~\citet{ingster09} for~\eqref{eqn:harmonic} demonstrates that the additional error incurred by using estimated eigenfunctions of the Laplacian, rather than directly using the Fourier series, results in the critical radius widening by only at most a factor of $(\log n)^{h(\beta,d)/2}$ from the minimax critical radius over $W^{1,2}(\mathcal{X};1)$.
+A comparison of Theorem~\ref{thm:sobolev_testing_rate} with the rates derived by~\citet{ingster09} for a test based on $T_{\mathrm{harm}}$ demonstrates that the additional error incurred by using estimated eigenfunctions of the Laplacian, rather than directly using the Fourier series, results in the critical radius widening by only at most a factor of $(\log n)^{h(\beta,d)/2}$ from the minimax critical radius over $W^{1,2}(\mathcal{X};1)$.
 This means our test based on $T_{\textrm{spec}}$ is almost \textit{as good as}, in a minimax sense, the best possible test over $W^{1,2}(\mathcal{X};1)$.
-As mentioned previously, the problem when $d \geq 4$ requires different assumptions on $f$ and is characterized by a different minimax rate.
-
-\subsubsection{Weighted Sobolev norms.}
-
-However, we would like to argue that in certain cases, $T_{\textrm{spec}}$ is in fact \textit{better} than competitors such as $T_{\textrm{harm}}$.
-While we do not yet have theoretical results along these lines, intuitively $T_{\textrm{spec}}$ should be particularly sensitive to functions $f$ which have smooth deviations from $f_0 = 0$ in high-density regions, but are potentially much wigglier in low-density regions.
-We therefore intend to (i) alter our function class $\mathcal{H}$ to be defined as the unit ball in a weighted Sobolev norm; precisely
-\begin{equation*}
-\mathcal{H}_{\Pbb} = \set{f \in L^2(\mathcal{X}): \sum_{\abs{\alpha} \leq s}\int\ \int_{\mathcal{X}} \abs{D^{\alpha}f(x)}^2 p^2(x)\,dx}
-\end{equation*}
-(ii) analyze the performance of $T_{\mathrm{spec}}$ over $\mathcal{H}_{\Pbb}$, and (iii) provide example distributions $\Pbb$ where $T_{\mathrm{spec}}$ outperforms reasonable competitors over $\mathcal{H}_{\Pbb}$ (See Figure~\ref{fig:fig3} for some empirical examples in the two-sample density testing case).
-
-\subsubsection{Higher-order derivative classes.}
-Theorem~\ref{thm:sobolev_testing_rate} covers only the $\{1,2\}$-Sobolev ball.
-In order to extend to higher-order Sobolev classes (i.e.
-the Sobolev $\{s,2\}$-balls where $s \geq 2$), we will likely need to consider kernels $K_r(u,v) = K\left(\frac{\norm{u - v}}{r}\right)$ for which
+As mentioned previously, when $d \geq 4$ the problem  requires different assumptions on $f$ and is characterized by a different minimax rate.
+
+We have shown that the test $\phi_{\mathrm{spec}}$ has some desirable theoretical properties.
+However, more comprehensive theoretical work is required to fully motivate $\phi_{\mathrm{spec}}$, and to better characterize its performance.
+
+\subsubsection{Proposed work.}
+
+Theorem~\ref{thm:sobolev_testing_rate} shows that the test $\phi_{\mathrm{spec}}$ is nearly-minimax optimal over $W_d^{1,2}(\mathcal{X};L)$.
+We propose to study the minimax properties of $\phi_{\mathrm{spec}}$ over two related Sobolev spaces, (i) the higher-order Sobolev balls $W_d^{s,2}(\mathcal{X};L)$, and (ii) density-weighted versions of Sobolev spaces.
+
+\paragraph{Higher-order Sobolev spaces.}
+In order to extend our results to hold with respect to the balls $W_d^{s,2}(\mathcal{X};L)$, we will likely need to consider kernels $K_r(u,v) = K\left(\frac{\norm{u - v}}{r}\right)$ for which
 \begin{equation}
@@
 as is typical in nonparametric smoothing problems \cite{tsybakov08}.
-Choosing kernels which satisfy~\eqref{eqn:higher_order_kernel} as opposed to $K_r(x,y) = \1(\norm{x - y} \leq r)$ allows for tighter control on the quadratic functional
+Choosing kernels which satisfy~\eqref{eqn:higher_order_kernel}--as opposed to our current choice of kernel $K_r(x,y) = \1(\norm{x - y} \leq r)$--allows for tighter control on the quadratic functional
 \begin{equation*}
@@
 this control is critical to obtain minimax rates, and otherwise the analysis when $s = 1$ versus $s > 1$ is not substantially different.
-We therefore intend to show that with an appropriate choice of kernel $K$, along with tuning parameters $\kappa$ and $r$, the test statistics $T_{\mathrm{spec}}$ and $T_{\mathrm{spec}}^{(2)}$ (the latter of which we define in the following section) result in minimax optimal tests over all Sobolev $\{s,2\}$-balls, in the regression and density testing problems respectively.
+We therefore \textbf{intend} to show for any Sobolev ball $W_d^{s,2}(\mathcal{X};L)$ such that $4s > d$, if the kernel $K$ satisfies~\eqref{eqn:higher_order_kernel} and the tuning parameters $\kappa$ and $r$ are selected appropriately, the test $\phi_{\mathrm{spec}}$ is minimax optimal.
+
+\paragraph{Weighted Sobolev norms.}
+
+The aforementioned completed and proposed work shows/would show that our test $\phi_{\mathrm{spec}}$ has minimax properties as good as any other minimax optimal test; for instance, a test based on the statistic $T_{\mathrm{harm}}$ defined in~\eqref{eqn:harmonic}.
+However, we would like to argue that in certain cases, $T_{\textrm{spec}}$ is in fact \textit{better} than competitors such as $T_{\textrm{harm}}$.
+We know from results in Section~\ref{subsec:neighborhood_graphs_spectra} that the eigenvectors $\set{v_i}$ converge towards eigenfunctions of a \emph{weighted} Laplacian operator.
+At a high-level, these eigenfunctions display inhomogeneous smoothness as a function of density, appearing much more wiggly in regions of low-density.
+
+The behavior of these eigenfunctions should affect the performance of $T_{\mathrm{spec}}$ in predictable ways.
+In particular, the test statistic $T_{\mathrm{spec}}$ should be more sensitive than $T_{\mathrm{harm}}$ and other test statistics which perform more classical smoothing to functions $f$ which have smooth deviations from $f_0 = 0$ in high-density regions, but are potentially much wigglier in low-density regions.
+We will therefore consider the performance of our test $\phi_{\mathrm{spec}}$ over the unit ball in a weighted Sobolev norm
+\begin{equation*}
+\mathcal{W}_d^{1,2}(\mathbb{P};\mathcal{X};L) := \set{f \in L^2(\mathcal{X}): \sum_{\abs{\alpha} \leq s}\int\ \int_{\mathcal{X}} \abs{D^{\alpha}f(x)}^2 p^2(x)\,dx \leq 1}.
+\end{equation*}
+We \textbf{intend} to upper bound the critical radius $\epsilon_n(\phi_{\mathrm{spec}};\mathcal{W}_d^{1,2}(\mathbb{P};\mathcal{X};L))$ as a function of $\mathbb{P}$ as well as $d$ and $n$, and exhibit choices of $\mathbb{P}$ for which $\phi_{\mathrm{spec}}$ will provably outperform competitors such as tests based on $T_{\mathrm{harm}}$ in the sense of having smaller worst-case risk.
+(See Figure~\ref{fig:fig3} for some empirical examples in the two-sample density testing case).
 
@@
 where as before $v_1,\ldots,v_n$ are eigenvectors of the Laplacian $L = L(X,r)$.
-
-While we do not yet have theoretical results characterizing the behavior of $T_{\mathrm{spec}}^{(2)}$, Figure~\ref{fig:fig3} shows some instances in which $T_{\mathrm{spec}}^{(2)}$ outperforms a range of common nonparametric tests.
+Figure~\ref{fig:fig3} shows some instances in which $T_{\mathrm{spec}}^{(2)}$ outperforms a range of common nonparametric tests.
 We see that the examples support our intuition; $T_{\textrm{spec}}^{(2)}$ is especially strong when the difference in densities $p - q$ oscillates at a higher frequency in regions where the mixture density $\frac{1}{2}(p + q)$ is small.
+While these examples empirically motivate the use of $T_{\mathrm{spec}}^{(2)}$, at present we still lack a proper theoretical characterization of its performance.
+This is what we propose to study.
 
@@
 
-We intend to (i) derive upper bounds on the worst-case risk of the test $\phi_{\mathrm{spec}}^{(2)} = \1(T_{\mathrm{spec}}^{(2)} \leq \tau)$ over the class $C_d^{s}(\mathcal{X};L)$, and prove that it is minimax optimal, (ii) investigate the minimax rate for two-sample density testing over the Sobolev class $\mathcal{W}_d^{s,2}(\mathcal{X};L)$, and (iii) prove an analogue to Theorem~\ref{thm:sobolev_testing_rate} establish that $\phi_{\mathrm{spec}}^{(2)}$ remains minimax optimal over the larger classes $\mathcal{W}_d^{s,2}(\mathcal{X};L)$.
-
-\subsubsection{Density testing for irregular alternatives.}
-As mentioned in Section~\ref{subsubsec:density_testing}, the regression and density testing models over e.g.
+\subsubsection{Proposed work.}
+
+As previously mentioned the minimax critical radius for the two-sample density testing problem over $C_d^{s}(\mathcal{X};L)$ is known to be $\epsilon_n^{\star}(C_d^{s}(\mathcal{X};L)) \asymp n^{-2s/(4s + d)}$ for all $d \geq 1$ and $s > 0$.
+Letting $\phi_{\mathrm{spec}}^{(2)} := \1\{T_{\mathrm{spec}} > \tau\}$, we \textbf{intend} to show that the critical radius $\epsilon_n(\phi_{\mathrm{spec}}^{(2)};C_d^{s}(\mathcal{X};L)) \asymp n^{-2s/(4s + d)}$ for all $d \geq 1, s > 0$, thus establishing that the test is minimax optimal over $C_d^{s}(\mathcal{X};L)$.
+
+There are certain technical nuances which make studying $T_{\mathrm{spec}}^{(2)}$ in the density testing problem more subtle than studying $T_{\mathrm{spec}}$ in the regression testing problem.
+(In particular, we require a condition on convergence of the eigenvectors ${v_i}$ which presently eludes us.) However, besides this issue by and large the technical results required are the same in each problem \emph{when the condition $4s > d$ is met}.
+
+\paragraph{Density testing for irregular ($4s \leq d$) alternatives.}
+
+% As mentioned in Section~\ref{subsubsec:density_testing}, the regression and density testing models over e.g.
 Holder spaces are formally asymptotically equivalent; however, these asymptotic equivalence statements hold only when $2s > d$.
 To illuminate some differences between the density and regression testing problems when the regression function $f$ or difference in densities $p - q$ is considered to be irregular (meaning $2s < d$), it is helpful to consider the following generative model: suppose we observe pairs $(x_i,\ell_i)$ for $i = 1,\ldots, (n + m)$, where as usual $x_i \sim \Pbb$ are independent, and additionally $\ell_i$ are conditionally independent given $x_i$ with distribution
-\begin{equation*}
-\ell_i = 
-\begin{cases*}
-1, & ~~  \textrm{with probability $\frac{p(x)}{p(x) + q(x)}$} \\
--1, & ~~ \textrm{with probability $\frac{q(x)}{p(x) + q(x)}$.} 
-\end{cases*}
-\end{equation*}
-We can relate this to the density testing problem by applying Bayes rule, whence we derive that the conditional distribution of $X_i|\ell_i = 1$ is $\mathbb{P}$, and similarly the conditional distribution of $X_i|\ell_i = -1$ is $\mathbb{Q}$.
+%\begin{equation*}
+%\begin{cases*}
+%1, & ~~  \textrm{with probability $\frac{p(x)}{p(x) + q(x)}$} \\
+%-1, & ~~ \textrm{with probability $\frac{q(x)}{p(x) + q(x)}$.} 
+%\end{cases*}
+%\end{equation*}
+%We can relate this to the density testing problem by applying Bayes rule, whence we derive that the conditional distribution of $X_i|\ell_i = 1$ is $\mathbb{P}$, and similarly the conditional distribution of $X_i|\ell_i = -1$ is $\mathbb{Q}$.
 On the other hand, since $\ell_i = f(x_i) + w_i$ where $f = \frac{p - q}{p + q}$ and $w_i$ is a mean-zero error term, this resembles in some respects the regression testing setup.
@@
 The fact that the variance of $w_i$ is unknown is crucial, and as previously mentioned it implies that there will no longer be a elbow in rates at $d = 4$.
-Indeed, as previously stated the minimax critical radius for the two-sample density testing problem over $C_d^1(\mathcal{X};L)$ is known to be $\epsilon_n^{\star}(C_d^1(\mathcal{X};L)) \asymp n^{-2/(4 + d)}$ for all $d$, including $d \geq 4$.
-
-When $d \geq 4$, choosing the hyperparameters for $T_{\mathrm{spec}}^{(2)}$ becomes more subtle.
-Statistics such as $T_{\mathrm{harm}}$ -- which take empirical evaluations of an orthogonal basis in $L^2(\mathcal{X})$--continue to be minimax optimal over $C_d^{1}(\mathcal{X};L)$ when $\kappa \asymp n^{2d/(4+d)}$.
-However, since this choice implies $\kappa > n$ the statistic $T_{\mathrm{spec}}^{(2)}$ -- which takes an orthogonal basis in $L^2(\mathbb{P}_n)$ -- reduces to simply $T_{\mathrm{spec}}^{(2)} = 1$ regardless of $\Pbb_n$ and $\Qbb_n$.
-Intuition from chi-squared tests--where the optimal choice of bin-width is on the order of $n^{-2/(4+d)}$, less than $n^{-1/d}$ whenever $d > 4$--suggests that we should modify our test statistic by choosing $r$ to be shrinking at a rate faster than $n^{-\frac{1}{d}}$.
+Indeed,
+
+Before precisely stating what we intend to show regarding $T_{\mathrm{spec}}^{(2)}$ when $4s \leq d$, it is worth emphasizing the differences between the density testing and regression testing problems when $4s \leq d$.
+In particular, in the density testing problem, the minimax critical radius is $\epsilon_n^{\star}(C_d^{s}(\mathcal{X};L)) \asymp n^{-2s/(4s + d)}$ for all $d \geq 1$ and $s > 0$.
+By contrast, the following series of inclusions:
+\begin{equation*}
+C_d^{s}(\mathcal{X};L) \subseteq \{f: \sup_{x \in \mathcal{X}}\abs{f(x)}\} \subseteq \{f: \norm{f}_{\mathcal{L}^4} \leq L\}
+\end{equation*}
+and the aforementioned work of \citet{guerre02} show that in the regression testing problem, the minimax critical radius is $\epsilon_n^{\star}(C_d^{s}(\mathcal{X};L)) = O(n^{-1/4})$.
+
+The upper bounds on the minimax critical radius in the density testing problem can be proved using binning tests, or tests based on truncated series test statistics such as $T_{\mathrm{harm}}$.
+These latter statistics--which take empirical evaluations of an orthonormal basis in $\mathcal{L}^2(\mathcal{X})$--are minimax optimal over $C_d^{1}(\mathcal{X};L)$ when the number of basis functions is chosen to be $\kappa \asymp n^{2d/(4s+d)}$, meaning $\kappa > n$ when $4s < d$.
+Since we have only $n$ eigenvectors $v_1,\ldots,v_n$, there is no clear relationship between $T_{\mathrm{harm}}$ and $T_{\mathrm{spec}}^{(2)}$ in this setting.
+
+We will therefore need a new approach to choosing $r$ and $\kappa$ if we want our test $\phi_{\mathrm{spec}}^{(2)}$ to be minimax optimal when $4s \leq d$.
+Intuition from binning tests--where the optimal choice of bin-width is on the order of $n^{-2s/(4s+d)}$, less than $n^{-1/d}$ whenever $4s < d$--suggests that we should modify our test statistic by choosing $r$ to be shrinking at a rate faster than $n^{-\frac{1}{d}}$.
 Of course at this connectivity scale with high probability the graph $G_{n,r}$ is disconnected \cite{penrose99}, and our analysis will necessarily be quite different than before.
-We intend to pursue such a modification and the resulting analysis.
+We \textbf{intend} to study the behavior of the graph $G_{n,r}$ when $r = o(n^{-1/d})$, with the hopes of showing that for $r \asymp n^{-2s/(4s+d)}$ and appropriate choices of $\kappa$ and $\tau$, the test $\phi_{\mathrm{spec}}^{(2)}$ has critical radius $\epsilon_n(\phi_{\mathrm{spec}}^{(2)}; C_d^{s}(\mathcal{X};L)) \asymp n^{-2s/(4s + d)}$ even when $4s \leq d$.
 
@@
 
-We define a pair of graph-based test statistics on the neighborhood graph $G_{n,r}$, which we cast as IPMs over different ``function'' classes $\Theta \subset \Reals^n$.
-
-\subsubsection{A graph-based Sobolev IPM.}
-
-We term our first test statistic the \emph{graph Sobolev} IPM, since the function class $\Theta$ over which we take the supremum will resemble a continuous Sobolev space. Formally we let
+In this section, we introduce a pair of graph-based test statistics on the neighborhood graph $G_{n,r}$, which we cast as IPMs over different ``function'' classes $\Theta \subset \Reals^n$.
+We have not substantively investigated the behavior of these test statistics yet, and therefore we have several lines of proposed work.
+
+\subsubsection{Proposed work: Sobolev IPM on a neighborhood graph.}
+
+We call our first proposed test statistic the \emph{graph Sobolev} IPM, since the function class $\Theta$ over which we take the supremum will resemble a continuous Sobolev space.
+Formally, we introduce the scaling $C_n := nr^{(d + 2)/2}$, and let
 \begin{equation}
 \label{eqn:sobolev_IPM}
-T_{\textrm{sob}} := \sup_{\theta \in \Reals^n: \norm{B^{(s)}\theta}_2 + \lambda\norm{\theta}_2 \leq 1} \abs{a^T \theta} = \sqrt{a^T (L^s + \lambda I)^{-1}a}, ~~ \lambda > 0, 
-\end{equation}
-where we recall the notation $a = \left(N^{-1},\ldots,N^{-1},-M^{-1},\ldots,-M^{-1}\right)$.
+T_{\textrm{sob}} := \sup_{\theta \in \Theta_{1,2}(\lambda)} \abs{\frac{1}{n}\sum_{i = 1}^{n} a_i \theta_i}, \quad  \Theta_{1,2}(\lambda) := \{\theta \in \Reals^n:~ C_n^{-1} \norm{B\theta}_2 + \frac{\lambda}{n^{1/2}}\norm{\theta}_2 \leq 1\} ~~ \lambda > 0,
+\end{equation}
+where we recall $B$ is the incidence matrix of $G_{n,r}$, and the notation $a = \left(N^{-1},\ldots,N^{-1},-M^{-1},\ldots,-M^{-1}\right)$.
 (We could easily adapt this statistic to the regression testing case by replacing $a$ with $y$, although we would lose the analogy to an IPM).
-
-\textbf{Remark:}~~Two aspects of the statistic $T_{\textrm{sob}}$ are worthy of comment.
-The first regards the role played by the parameter $\lambda$, which is at first glance somewhat mysterious.
-It may seem more natural to consider the statistic
-\begin{equation*}
-\sup_{\theta \in \Reals^n: \norm{B\theta}_2 \leq 1} \abs{a^T \theta}, 
+Several aspects of the statistic $T_{\textrm{sob}}$ are worthy of comment.
+
+\textbf{Remark:}~~The scalings $C_n$ and $n^{-1/2}$ in front of $\norm{B\theta}_2$ and $\norm{\theta}_2$ are the proper scalings at which to compare these norms to the continuum norms $\norm{\cdot}_{W_d^{s,2}(\mathcal{X})}$ and $\norm{\cdot}_{\mathcal{L}^2(\mathcal{X})}$.
+These scalings are strictly for convenience, as we could always absorb them into the parameter $\lambda$.
+
+\textbf{Remark:} The role played by the term $\frac{\lambda}{n^{1/2}}\norm{\theta}_2$ is at first glance somewhat mysterious. It may seem more natural to consider the statistic
+\begin{equation*}
+\sup_{\theta \in \Reals^n: C_n^{-1}\norm{B\theta}_2 \leq 1} \abs{\frac{1}{n}\sum_{i = 1}^{n} a_i \theta_i}, 
 \end{equation*}
 so that the function class is defined only with respect to the semi-norm $\norm{B\theta}_2$.
-However, it can be shown that this function class is too rich, and our test statistic will overfit to noise; hence we restrict it in the manner of~\eqref{eqn:sobolev_IPM}, a modification similar to that of \citet{arbel18} who observe that it improves empirical performance.
-Second, we note that in general, an analogous IPM in the continuum Sobolev space is a poor candidate for testing.
-Specifically, we have that when $2s < d$,
+However, we have shown in preliminary calculations that this function class is too rich, and that the resulting test statistic will overfit to noise; hence we restrict it in the manner of~\eqref{eqn:sobolev_IPM}, a modification similar to that of \citet{arbel18} who observe that it improves empirical performance.
+
+\textbf{Remark:} Finally we note that in general, an analogous IPM in the continuum Sobolev space is a poor candidate for testing. Specifically, we have that when $2s < d$,
 \begin{equation}
 \label{eqn:continuum_Sobolev_IPM}
-\sup_{f \in L^2(\mathcal{X}): \norm{f}_{W^{s,2}(\mathcal{X})} \leq L} \abs{\mathbb{P}_nf - \mathbb{Q}_nf} = \infty
+\sup_{f \in W_d^{s,2}(\mathcal{X};1)} \abs{\mathbb{P}_nf - \mathbb{Q}_nf} = \infty
 \end{equation}
@@
 
-Based on the success of Sobolev IPMs in real-world problems \citet{mroueh17,arbel18} and the favorable properties of analogous estimators when performing regression on a grid \citet{sadhanala16}, we believe testing with $T_{\textrm{sob}}$ is a reasonable idea.
-We intend to investigate the minimax properties of such a test with respect to the function classes $C^s(\mathcal{X})$ and $W^{s,2}(\mathcal{X})$.
-
-\subsubsection{A graph-based Total Variation (TV) IPM.}
+Based on the success of Sobolev IPMs\footnote{that is, IPMs such as~\eqref{eqn:continuum_Sobolev_IPM} but where $2s > d$ or some other modification has been made to avoid their being ill-posed} in real-world problems \citet{mroueh17,arbel18} and the favorable properties of analogous estimators when performing regression on a grid \citet{sadhanala16}, we believe testing with $T_{\textrm{sob}}$ is a reasonable idea.
+We will investigate the matter theoretically and empirically.
+On the theoretical side, we \textbf{intend} to show that the test  $\phi_{\mathrm{sob}} := \1\{T_{\mathrm{sob}} \leq \tau\}$ achieves minimax optimal density testing radius $\epsilon_n(\phi_{\mathrm{sob}};W_d^{s,2}(\mathcal{X};L)) \asymp n^{-2s/(4s + d)}$.
+On the empirical side, we note that $T_{\mathrm{spec}}^{(2)}$ and $T_{\mathrm{sob}}$ appear at a high level to operate  by regularizing the spectrum of $L$, albeit in different manners; this makes it unclear when and why we should prefer one test over the other.
+We \textbf{intend} to elucidate these differences, and motivate the need for both tests, by showing directions (that is, specific choices of $p$ and $q$ in the alternative) for which $T_{\mathrm{spec}}^{(2)}$ outperforms $T_{\mathrm{sob}}$ and vice versa.
+
+\subsubsection{Proposed work: TV IPM on a neighborhood graph.}
 Our second IPM test statistic will resemble the first, except we replace the graph Sobolev norm by a graph total variation norm. We hence call this statistic the \emph{graph TV} IPM, defined as
 \begin{equation*}
-T_{\mathrm{TV}} := \sup_{\theta \in \Reals^n: \norm{B\theta}_1 + \lambda\norm{\theta}_2 \leq 1} \abs{a^T \theta}.
-\end{equation*}
-To motivate the graph TV IPM, we give some background on testing in TV spaces.
-
-\paragraph{Testing in TV spaces.}
-
-To one extent or another, functions in the Sobolev and Holder balls exhibit homogeneous smoothness.
-In reality we may not wish to restrict our attention to such functions.
-This is in part the motivation behind considering functions of bounded variation.
-Formally, for a function $f \in L^1(\mathcal{X})$ the \emph{total variation} semi-norm of $f$ is \citet{evans15}
-\begin{equation*}
-TV(f;\mathcal{X}) := \sup \left\{ \int_U f \, \dive \, \psi \,dx : \psi \in C_c^1(\mathcal{X}; \Reals^d), \abs{\psi} \leq 1 \right\};
-\end{equation*}
-and we write $BV_d(\mathcal{X})$ for the subset of functions $f \in L^1(\mathcal{X})$ which have bounded norm
-\begin{equation*}
-\norm{f}_{BV_d(\mathcal{X})} := \norm{f}_{\infty} + TV(f;\mathcal{X}).
-\end{equation*}
-
-In the one-dimensional case, some work has been done on nonparametric testing over the Besov spaces $B_{p,q}^s(\mathcal{X})$ (for a formal definition see Section 4.3 of \cite{gine16}), which we can relate to bounded variation spaces via the embeddings
-\begin{equation}
-\label{eqn:besov_embeddings}
-B_{11}^1(\mathcal{X}) \subseteq BV_1(\mathcal{X}) \subseteq B_{1\infty}^1(\mathcal{X}).
-\end{equation}
-
-\citet{lepski1999} study the regression testing problem, with data observed according to the white noise model as in \eqref{eqn:gaussian_white_noise}.
-They derive among other things that for the Besov balls $B_{1,q}^s(\mathcal{X}; L)$, the minimax critical radius $\epsilon_n^{\star}(B_{1,q}^s(\mathcal{X}; L)) \asymp n^{-2s'/(4s' + 1)}$, where $s' = s - \frac{1}{4}$ and the result holds for any $s > 1, q \geq 1$.
-\citet{ingster00} improve on this by showing the same result holds whenever $s > 1/2$.
-This along with~\eqref{eqn:besov_embeddings} implies that $\epsilon_n^{\star}(BV_1(\mathcal{X})) \asymp n^{-\frac{3}{8}}$.
-
-When $d \geq 2$ minimax testing rates over the Besov and bounded variation spaces are still unknown.
-However, study of the equivalent estimation problem suggests the answers may be substantially different from the 1d problem.
-\citet{sadhanala16} consider a discretized version of this problem -- where observations are made on a discrete grid and the regression function is assumed to lie within a discrete bounded variation class -- and show that the minimax estimation rate in empirical norm is on the order of $n^{-1/d}$ for $d \geq 2$.
-This is slower than minimax estimation rate over the Sobolev space $\mathcal{W}_d^{1,2}$, the largest Sobolev space inscribed within $BV_d$.
-\citet{delyon96,kerkyacharian08,lepski2015,ruiz18} all consider the regression white noise model \eqref{eqn:gaussian_white_noise} and show, in progressive degrees of generality, that the minimax rates over the Besov spaces $B_{p,q}^s$ exhibit a phase transition at the boundary $2s = d$.
-When $2s > d$, the minimax estimation rate (with loss measured in squared $\mathcal{L}_2$ norm) is the standard $n^{-2s/(2s + d)}$; on the other hand when $2s < d$, the minimax rate is $n^{-s/d}$ which is much slower.
-When $s = 1$, this phase transition occurs at $d = 2$, and so in a minimax sense the univariate and multivariate estimation problems over $BV_d$ are revealed to be quite different.
-\citet{sadhanala17} show that this same phenomenon carries over to the grid, and the discrete TV classes defined upon it.
-
-We intend to investigate whether such a phase transition (from $d = 1$ to $d \geq 2$) exists in the testing setting as well, by deriving upper and lower bounds for the critical radius $\epsilon_n(BV_d(\mathcal{X};L))$, where $BV_d(\mathcal{X}; L) = \set{f \in BV_d(\mathcal{X}): \norm{f}_{BV_d(\mathcal{X})} \leq L}$.
-Our upper bound will be based on the test $\phi_{TV} := \1\{T_{\mathrm{TV}} \geq \tau \}$.
-
-\subsubsection{Computational considerations.}
+T_{\mathrm{TV}} := \sup_{\theta \in \Theta_{1,1}(\lambda)} \abs{\frac{1}{n}\sum_{i = 1}^{n} a_i \theta_i}, \quad \Theta_{1,1}(\lambda) := \{\theta \in \Reals^n:~ (C_n')^{-1} \norm{B\theta}_1 + \frac{\lambda}{n^{1/2}}\norm{\theta}_2 \leq 1\}
+\end{equation*}
+where $C_n' := n^{2}r^{(d + 1)}$.
+Compared to the Sobolev case, far less methodological and theoretical work has been done regarding total variation-based test statistics.
+As previously mentioned, the minimax lower bound on the critical radius over the bounded variation spaces remains unknown for the multivariate problem, even under idealized models such as~\eqref{eqn:gaussian_white_noise}.
+We \textbf{intend} to investigate whether a phase transition such as that which occurs in the analogous estimation problem (from $d = 1$ to $d \geq 2$) exists in the testing setting as well, by deriving upper and lower bounds for the critical radius $\epsilon_n(BV_d(\mathcal{X};L))$ where $BV_d(\mathcal{X}; L) = \set{f \in BV_d(\mathcal{X}): \norm{f}_{BV_d(\mathcal{X})} \leq L}$.
+Our upper bound will be based on the test $\phi_{TV} := \1\{T_{\mathrm{TV}} \geq \tau \}$, on the theory that the test statistic $T_{\mathrm{TV}}$ is a good approximation to the functional
+\begin{equation*}
+\sup_{g:~ \mathrm{TV}(g;\mathcal{X}) + \lambda\norm{g}_{\mathcal{L}^2} \leq 1} \abs{\int_{\mathcal{X}} g(x) \,d\Pbb(x) - \int_{\mathcal{X}} g(x) \,d\Qbb(x)}
+\end{equation*}
+and is therefore a reasonable choice for testing over $BV_d(\mathcal{X};L)$.
+
+\subsection{Proposed work: computational considerations.}
 
@@
 
-When $n$ is large and the graph $G_{n,r}$ is relatively dense, exactly computing the eigenvectors of $L$ (as in~\eqref{eqn:graph_spectral_projections}) or the matrix inverse $(L + \lambda I)^{-1}$ (as in ~\eqref{eqn:sobolev_IPM}) may be computationally expensive.
+When $n$ is large and the graph $G_{n,r}$ is relatively dense, exactly computing the eigenvectors of $L$ (as in~\eqref{eqn:graph_spectral_projections}) or the matrix inverse $(L + \lambda I)^{-1}$ (implicitly in \eqref{eqn:sobolev_IPM}) may be computationally expensive.
 Recently, there has been remarkable progress made (e.g.
-\citet{spielman2011,batson13,spielman2014,cohen14}) in efficiently computing $\epsilon$-\emph{spectral sparsifiers} of $G$: that is, graphs $\widetilde{G}$ with Laplacian matrices $\widetilde{L}$ which satisfy:
+\citet{spielman2011,batson13,spielman2014,cohen14}) in efficiently computing $\delta$-\emph{spectral sparsifiers} of $G$: that is, graphs $\widetilde{G}$ with Laplacian matrices $\widetilde{L}$ which satisfy:
 \begin{equation}
 \label{eqn:spectral_similarity}
-(1 - \epsilon)x^T \widetilde{L} x \leq x^T L x \leq (1 + \epsilon) x^T \widetilde{L} x, \quad \textrm{for some $0 \leq \epsilon < 1$, and all $x \in \Reals^n$.}
-\end{equation}
-and have many fewer edges than $G$ (hence the term ``sparsifier''.) Since $\widetilde{L}$ is a sparse matrix, we can compute its eigenvectors $\set{\widetilde{v}}$ or the inverse $(\widetilde{L} + \lambda I)^{-1}$  efficiently.
-Additionally \eqref{eqn:spectral_similarity} should imply that the resulting test statistics behave similarly to those defined with respect to $L$.
+(1 - \delta)x^T \widetilde{L} x \leq x^T L x \leq (1 + \delta) x^T \widetilde{L} x, \quad \textrm{for some $0 \leq \delta < 1$, and all $x \in \Reals^n$,}
+\end{equation}
+and have many fewer edges than $G$ (hence the term ``sparsifier''.) Since $\widetilde{L}$ is a sparse matrix, we can compute its eigenvectors $\set{\widetilde{v}_i}$ or the inverse $(\widetilde{L} + \lambda I)^{-1}$  efficiently.
+Additionally, \eqref{eqn:spectral_similarity} should imply that the resulting test statistics behave similarly to those defined with respect to $L$.
 As yet limited work \citet{vonluxburg2014,sadhanala16b} has been done to rigorously analyze the performance of spectral sparsifiers in statistical learning problems, despite their obvious promise.
-We plan to address this gap by providing a theoretical characterization of the performance of graph statistics on sparsified graphs in the regression and density testing problems.
-
-% We also mention in passing that another potential computational difficulty stems from calibrating the threshold $\tau$.
+We \textbf{intend} to address this gap by showing how graph statistics computed on sparsified graphs perform in the regression and density testing problems.
+Specifically, we hope to show that the statistics $\widetilde{T}_{\mathrm{spec}}, \widetilde{T}_{\mathrm{spec}}^{(2)}$ and $\widetilde{T}_{\mathrm{sob}}$ -- equivalent to their \textcolor{red}{non-tilde} counterparts but computed over the sparsified $\widetilde{L}$ rather than the original matrix $L$ -- have critical radii on the same order as those of their \textcolor{red}{non-tilde} counterparts, while being much easier to compute.
+
+We also mention in passing that another potential computational difficulty stems from calibrating the threshold $\tau$.
 While the choice of $\tau$ in the statement of Theorem~\ref{thm:sobolev_testing_rate} is optimal in a minimax sense, practically speaking depending on the regression function $f$ and distribution $\Pbb$ it may be a conservative threshold, with a resulting loss in power.
 In practice, one can calibrate $\tau$ by permutation, however doing so requires repeatedly shuffling the data $X$ and recomputing a test statistic $T$.
-It would therefore be helpful to know the distribution of our various proposed statistics $T$ under the null hypothesis, or at least derive asymptotic approximations to these distribution, and we plan to investigate this.
-However, while in the density testing context there has been some work on deriving asymptotic distributions of IPMs \cite{gretton12,sadhanala19}, typically these limiting distributions depend on the unknown $\mathbb{P}$, and therefore the practical implications of these results is somewhat limited.
+It would therefore be helpful to know the distribution of our various proposed statistics $T$ under the null hypothesis, or at least derive asymptotic approximations to these distributions.
+In the density testing context there has been some work on deriving asymptotic distributions of IPMs \cite{gretton12,sadhanala19}, but typically these limiting distributions depend on the unknown $\mathbb{P}$, and therefore the practical implications of these results is somewhat limited.
+Nevertheless, we \textbf{intend} to derive asymptotic null distributions for the statistics we propose, and investigate whether they indeed depend on $\mathbb{P}$.
+
+\subsection{Estimated timeline.}
+
+The following is an estimated timeline for completing my proposed work.
+\begin{itemize}
+	\item \textbf{Spring 2020:} Complete all aforementioned intended work regarding the tests $\phi_{\mathrm{spec}}$ and $\phi_{\mathrm{spec}}^{(2)}$ 
+	\item \textbf{Fall 2020:} Complete all aforementioned intended work regarding the test $\phi_{\mathrm{sob}}$, and address computational considerations.
+	\item \textbf{Spring 2021:} Complete all aforementioned intended work regarding the test $\phi_{\mathrm{TV}}$, and minimax testing over the bounded variation spaces $BV_d(\mathcal{X})$.
+Defend my thesis.
+\end{itemize}
 

### 2019-11-23 notes/weekly_notes/11.15.19.NotesForWeek.tex
added weekly notes

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\domain}{\mathcal{X}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\Partial}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 11/16/19 - 11/23/19}
\author{Alden Green}
\date{\today}
\maketitle

In this week's notes, I pick up the thread of the ``11.8.19.NotesForWeek'' document. There, we obtained that the critical radius in the two-sample density testing problem over a graph $G$ was a function of
\begin{equation*}
\Pi_{\max} = \max_{i = 1,\ldots,n} \set{\sum_{k = 1}^{\kappa} v_{k,i}^2}
\end{equation*}
where $L = V S V^T$ and $V = (v_1,\ldots,v_n)$ is an orthonormal matrix, $S$ is a diagonal matrix with increasing entries $0 = s_1 \leq s_2 \leq \ldots \leq s_n$. In particular, we showed that if $\Pi_{\max}$ could be upper bounded by a constant less than $1$, then the critical radius in the two-sample density testing problem over a graph $G$ was the same as in the one-sample regression testing problem. Our goal was therefore to upper bound $\Pi_{\max} \leq \frac{1}{2}$ for an appropriate choice of $\kappa$. We showed that such an upper bound holds when $G = G_{dn}$ the $d$-dimensional grid, and $\kappa \leq \frac{n}{4}$.

\section{Bounding $\Pi_{\max}$ for neighborhood graphs in $1$-dimension.}
Now, suppose we additionally have a graph $\widetilde{G}$ with Laplacian matrix $\widetilde{L}$ such that
\begin{equation}
\label{eqn:spectral_similarity}
(1 - \delta) x^T L x \leq x^T \widetilde{L} x \leq (1 + \delta) x^T L x, \quad \textrm{for some $0 \leq \delta \leq 1$ and all $x \in \Reals^n$.}
\end{equation}
Write the spectral decomposition $\widetilde{L} = U \Lambda U^T$, and suppose that the eigenvectors $U$ satisfy
\begin{equation}
\label{eqn:incoherence_tilde}
u_{l,i}^2 \leq \frac{2}{n} \quad \textrm{for all $l,i = 1,\ldots,n$.}
\end{equation}

\begin{lemma}
	\label{lem:incoherence}
	Let $\widetilde{G} = G_{1n}$ be the 1-dimensional grid graph (i.e. the chain), and assume $G$ satisfies \eqref{eqn:spectral_similarity}. Then we have the following bound on incoherence of the eigenvectors $v_{k,i}$: for any $R \gtrsim n \delta^{1/2} (s_{\max}^{1/2} \vee 1)$,
	\begin{equation*}
	v_{k,i}^2 \lesssim \frac{1}{n}\left(R + \frac{n^4 \delta^2 s_{\max}^2}{R^3}\right) \quad \textrm{for any $i = 1,\ldots,n$}
	\end{equation*} 
	for any $k \leq n/8$.
\end{lemma}

Before proving Lemma~\ref{lem:incoherence}, we demonstrate its consequence. To state things concisely, we use tilde-order notation $\widetilde{O}$ and $\widetilde{\Theta}$ to suppress $\log(n)$ terms. Assume that 
\begin{equation}
\label{eqn:neighborhood_graph_spectral_similarity}
\textrm{$G = G_{n,r}$, $\widetilde{G} = G_{1n}$ satisfy \eqref{eqn:spectral_similarity} with $\delta = \widetilde{O}(n^{-1})$}, \quad \textrm{the maximum degree $d_{\max}(G) = \widetilde{O}(1)$}
\end{equation}
Using the upper bound on maximum eigenvalue of a Laplacian $s_{\max} \leq 2 d_{\max}$), by Lemma~\ref{lem:incoherence} the eigenvectors $v$ of $G$ satisfy
\begin{equation*}
v_{k,i}^2 = \widetilde{O}\biggl(\frac{1}{n}\left(R + \frac{n}{R^3}\right)\biggr)
\end{equation*}
for any $k \leq n/8$ and $R = \widetilde{\Theta}(n^{1/2}s_{\max}^{1/2}) = \widetilde{\Theta}(n^{1/2})$. Taking $R = \widetilde{\Theta}(n^{1/2})$, we have that
\begin{equation}
\label{eqn:incoherence}
\max_{i = 1,\ldots,n} v_{k,i}^2 = \widetilde{O}\left(n^{-1/2}\right)
\end{equation}
for all $k \leq n/8$. This is quite a bit weaker than the incoherence bound~\eqref{eqn:incoherence_tilde} that the eigenvectors of the chain satisfy. Nevertheless, it will suffice for our purposes. When $\kappa \asymp n^{2/5}$, equation~\eqref{eqn:incoherence} implies that
\begin{equation*}
\Pi_{\max} = \widetilde{O}(\kappa n^{-1/2}) = o(1). 
\end{equation*}
We have previously shown that when $r = \widetilde{O}(n^{-1})$, the graph $G_{n,r}$ satisfies~\eqref{eqn:neighborhood_graph_spectral_similarity}  with high probability; therefore, $\Pi_{\max}$ is less than constant order. Looking back at the ``11.8.19.NotesForWeek'' (specifically equation (8) which gives the critical radius) we see that this is sufficient to ensure that the critical radius in the two-sample density testing and one-sample regression testing problems will be the same when $d = 1$. 
\section{Technical Results.}

\subsection{Proof of Lemma~\ref{lem:incoherence}}

The proof of Lemma~\ref{lem:incoherence} will proceed according to the following steps.
\begin{enumerate}
	\item We show that the incoherence $\max_{i} v_{k,i}^2$ can be upper bounded
	\begin{equation}
	\label{eqn:incoherence_proof_1}
	\max_{i} v_{k,i}^2 \leq \frac{1}{n} \left(\sum_{j = 1}^{n} \abs{\dotp{v_k}{u_j}}\right)^2
	\end{equation}
	replacing the norm $\norm{v_k}_{\infty}$ by the inner products $\dotp{v_k}{u_j}$.
	\item Let
	\begin{equation*}
	I(k,r) = \set{j \geq 0: \abs{j - k} \leq r}.
	\end{equation*}
	Using Davis-Kahan, we establish that the following upper bounds
	\begin{equation}
	\label{eqn:incoherence_proof_2}
	\sum_{j \not\in I(k,r)}^{n} (\dotp{v_k}{u_j})^2 \leq 400 \frac{n^4 \delta^2 s_{\max}^2}{R^4}
	\end{equation}
	hold for each $k \leq n/8$ and any $r \geq 8k\sqrt{\delta}$. 
	\item We carefully upper bound the $L_1$ norm present in \eqref{eqn:incoherence_proof_1} given the various bounds on $L_2$ norm we've established in \eqref{eqn:incoherence_proof_2}.
\end{enumerate}

\paragraph{Step 1: Proof of~\eqref{eqn:incoherence_proof_1}.}

We can re-express $v_k$ in the basis $(u_j)$ as
\begin{equation*}
v_k = \sum_{j = 1}^{n} \dotp{v_k}{u_j} u_j,
\end{equation*}
whereupon the bound~\eqref{eqn:incoherence_proof_1} follows from \eqref{eqn:incoherence_tilde}. 
\paragraph{Step 2: Proof of~\eqref{eqn:incoherence_proof_2}.}

Let $\widetilde{L} = L + H$. By Davis Kahan, we have that
\begin{equation}
\label{eqn:davis_kahan}
\sum_{j \not\in I(k,r)} (\dotp{v_k}{u_j})^2 \leq \left(\frac{\norm{H}_{\textrm{op}}}{\min{\abs{\lambda_j - s_k}:j \in I(k,r)}}\right)^2
\end{equation}
To upper bound the numerator, we use~\eqref{eqn:spectral_similarity} to get
\begin{equation*}
\norm{H}_{op} \leq \delta s_{\max}.
\end{equation*}

To lower bound the denominator, we note
\begin{align}
\abs{\lambda_j - s_k} & \geq \abs{\lambda_j - \lambda_k} - \abs{\lambda_k - s_k} \nonumber \\
& \geq \abs{\lambda_j - \lambda_k} - \delta \lambda_k. \label{eqn:incoherence_proof_8}
\end{align}
The eigenvalues of the chain graph are well known to be
\begin{equation*}
\lambda_j = 2\left(1 - \cos\left(\frac{j\pi}{n}\right)\right) = 4\sin^2\left(\frac{k\pi}{n}\right) ~~\textrm{for $k = 0,\ldots,n - 1$.}
\end{equation*}
By Taylor expansion, for $k \leq j \leq n/(2\pi)$ we have
\begin{equation}
\label{eqn:incoherence_proof_5}
\lambda_j - \lambda_k = 2\left(\cos\left(\frac{k\pi}{n}\right) - \cos\left(\frac{j\pi}{n}\right)\right) \geq \frac{(k - j)^2\pi^2}{n^2}.
\end{equation}
and when $j \geq n/(2\pi)$ we have $\lambda_j - \lambda_k \geq 2(\cos\left(\frac{\pi}{8}\right) - \cos\left(\frac{1}{2}\right) > .09$. Additionally since $\sin(x) \leq x$ for all $x \geq 0$, we have
\begin{equation}
\label{eqn:incoherence_proof_6}
\lambda_k \leq 4\frac{k^2\pi^2}{n^2}.
\end{equation}
Combining~\eqref{eqn:incoherence_proof_5},\eqref{eqn:incoherence_proof_6}, and the lower bound $\abs{k - j} \geq R \geq 8k\sqrt{\delta}$, we have that
\begin{equation}
\label{eqn:incoherence_proof_7}
\abs{\lambda_j - \lambda_k} - \delta \lambda_k \geq .04\frac{(k - j)^2\pi^2}{n^2} - 4 \delta\frac{k^2\pi^2}{n^2} \geq .02\frac{(k - j)^2\pi^2}{n^2}.
\end{equation}
The result follows from \eqref{eqn:incoherence_proof_7}, \eqref{eqn:incoherence_proof_8} and \eqref{eqn:davis_kahan}.

\paragraph{Step 3: $L_1$ norm to $L_2$ norm.}

Let $a \in \Reals^n$, and suppose we know that the $L_2$ norm of $a$ is bounded,
\begin{equation}
\label{eqn:incoherence_proof_3}
\norm{a}_2^2 \leq 1.
\end{equation}
Under no other conditions on $a$, the upper bound $\norm{a}_1 \leq \sqrt{n}$ is the best that can be hoped for (achieved when $a = (n^{-1/2},\ldots,n^{-1/2})$). However, suppose we also know that for some $B_2^2 \geq B_3^2 \geq \ldots \geq B_n^2$, we have that there exists some $R \in [n]$ such that
\begin{equation}
\label{eqn:incoherence_proof_4}
\sum_{j = r}^{n} a_j^2 \leq B_r^2 ~~\textrm{for each $r = R,\ldots,n$.}
\end{equation}
Clearly, if $B_n^2 \leq \frac{1}{n}$ then $a$ cannot be equal to $(n^{-1/2},\ldots,n^{-1/2})$. We might hope for more general improvements on the bound $\norm{a}_1 \leq \sqrt{n}$ if $B_R^2$ is quite small once $R$ gets sufficiently large. The following Lemma gives such a result.
\begin{lemma}
	\label{lem:incoherence_util_1}
	Suppose $a \in \Reals^n$ satisfies \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4} for some $1 = B_1^2 = B_2^2 \geq \ldots \geq B_n^2 \geq B_{n+1}^2 = 0$. Assume additionally that there exists an $R \in [n]$ such that
	\begin{equation}
	\label{eqn:incoherence_util_1}
	B_r^2 - B_{r + 1}^2 \leq \frac{1}{r}~\textrm{for each $r = R, R+1,\ldots, n - 1$.}
	\end{equation}
	Then,
	\begin{equation}
	\label{eqn:incoherence_util_2}
	\norm{a}_1 \leq \sqrt{R}\sqrt{1 - B_R^2} + \sum_{j = R}^{n} \sqrt{B_j^2 - B_{j+1}^2}.
	\end{equation}
\end{lemma}
\begin{proof}
	Set
	\begin{equation*}
	(a_\star)_j^2 = 
	\begin{cases*}
	\frac{1}{R}(1 - B_R^2), ~ j \leq R \\
	B_j^2 - B_{j + 1}^2, j > R
	\end{cases*}
	\end{equation*}
	so that $\norm{a_\star}_1$ is equal to the right hand side of~\eqref{eqn:incoherence_util_2}. We now prove by contradiction that $a_\star$ maximizes $\norm{a}_1$ under the constraints \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4}. Suppose this were not true, and let $a$ be a vector such that $a$ satisfies \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4} and $\norm{a}_1 > \norm{a_{\star}}$. Then,
	\begin{itemize}
		\item since $\norm{a}_1 > \norm{a_{\star}}$ then there exists some index $j$ such that
		\begin{equation}
		\label{eqn:incoherence_util_3}
		a_j > (a_{\star})_j
		\end{equation} 
		\item by~\eqref{eqn:incoherence_proof_3} $\norm{a}_2 \leq 1$. Since $\norm{a_{\star}}_2 = 1$, by \eqref{eqn:incoherence_util_3} there must exist some index $k$ such that
		\begin{equation}
		\label{eqn:incoherence_util_4}
		(a_{\star})_k > a_k.
		\end{equation} 
		Choose $k$ to be the largest of all such indices.
		\item suppose $k < j$ and $j > R$. Since $k$ was chosen to be the largest such index which satisfied~\eqref{eqn:incoherence_util_4}, clearly $a_i \geq (a_{\star})_i$ for all $i > j$, and by \eqref{eqn:incoherence_util_3} $a_j > (a_{\star})_j$. But then 
		\begin{equation*}
		B_j^2 = \sum_{i = j}^{N} (a_{\star})_i^2 < \sum_{i = j}^{N} a_i^2 
		\end{equation*}
		and so $a$ does not satisfy~\eqref{eqn:incoherence_proof_4}. Therefore either $k > j$ or $j \leq R$.
		\item In either case, we have
		\begin{equation*}
		a_j > (a_{\star})_j \geq (a_{\star})_k > a_k.
		\end{equation*}
		Let $j \leq l < k$ be the largest index for which $a_j > (a_{\star})_j$.

		We now construct a vector $\norm{b}$ which satisfies the constraints \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4} such that $\norm{b}_1 > \norm{a}_1$. Once we have shown this, we will have established a contradiction, and the proof of Lemma~\ref{lem:incoherence_util_1} will be complete. Let $b = (b_i)$ be given as follows:
		\begin{equation*}
		b_i = 
		\begin{cases*}
		a_i, ~~\textrm{if $i \neq k,l$}, \\
		a_{\star,k}, ~~\textrm{if $i = k$}, \\
		\sqrt{a_l^2 - (a_{\star,k}^2 - a_k^2)}, ~~\textrm{if $i = l$.}
		\end{cases*}
		\end{equation*}
		Clearly $\norm{b}_2 = \norm{a}_2 = 1$ and so $b$ satisfies~\eqref{eqn:incoherence_proof_3}. To see that $b$ satisfies~\eqref{eqn:incoherence_proof_4} we separate the analysis into cases. When $r \geq l + 1$, we have that
		\begin{equation*}
		\sum_{i = r}^{n} b_i^2 = \sum_{i = r}^{n} a_i^2 \leq B_r^2.
		\end{equation*}
		When $l < r \leq k$, we have that
		\begin{equation*}
		\sum_{i = r}^{n} b_i^2 \leq \sum_{i = r}^{k} a_{\star,i}^2 + \sum_{i = k +1}^{n} a_{i}^2 \leq B_r^2 - B_{k + 1}^2 + B_{k + 1}^2 = B_r^2
		\end{equation*}
		Finally, when $R \leq r \leq l$, we have
		\begin{equation*}
		\sum_{i = r}^{n} b_i^2 \leq a_l^2 - (a_{\star,k}^2 - a_k^2) + (a_{\star,k}^2) + \sum_{i \neq k,l} a_i^2 = \sum_{i = r}^{n} a_i^2 \leq B_r^2.
		\end{equation*}
		Therefore $b$ satisfies~\eqref{eqn:incoherence_proof_4}. Finally, we have that
		\begin{equation*}
		\norm{b}_1 = \norm{a}_1 + (a_{\star,k}  - a_k) - (\sqrt{a_l^2 - (a_{\star,k}^2 - a_k^2)} - a_l) > \norm{a}_1 + (a_{\star,k}  - a_k) + (a_{\star,k}  - a_k) - (a_{\star,k}  - a_k) = \norm{a}_1,
		\end{equation*}
		so we have established the desired contradiction.
	\end{itemize}
\end{proof}

\paragraph{Putting the pieces together.}
We apply Lemma~\ref{lem:incoherence_util_1} to the vector $a = (a_i)_{i = 1}^{n}$, where
\begin{equation*}
a_i = \frac{1}{\sqrt{2}}\left(\abs{\dotp{v_k}{u_{k - i}}} + \abs{\dotp{v_k}{u_{k + i}}}\right).
\end{equation*}
Note the following:
\begin{itemize}
	\item $\norm{a}_2 \leq 1$.
	\item By \eqref{eqn:incoherence_proof_2}, the vector $a$ satisfies the constraint~\eqref{eqn:incoherence_proof_4} with
	\begin{equation*}
	B_r^2 = 400\frac{n^4\delta^2s_{\max}^2}{r^4} ~~\textrm{when $r \geq 8 k \sqrt{\delta}$.}
	\end{equation*}
	\item Taylor expanding $f(r + 1) = (r + 1)^{-4}$ around $r$, we have that
	\begin{equation*}
	B_r^2 - B_{r + 1}^2 \leq 3200 \frac{n^4 \delta^2 s_{\max}^2}{r^5} ~~\textrm{when $r \geq 2$.}
	\end{equation*}
	Therefore, the sequence $(B_r)$ satisfies \eqref{eqn:incoherence_util_1} for all $r$ large enough such that 
	\begin{equation*}
	r \geq 2^{5/4} 100^{1/4} n \sqrt{\delta s_{\max}}
	\end{equation*}
	\item Since $a$ and $(B_r)$ satisfy the constraints \eqref{eqn:incoherence_proof_3}, \eqref{eqn:incoherence_proof_4} and \eqref{eqn:incoherence_proof_5}, we may apply Lemma~\ref{lem:incoherence_util_1} to $a$, obtaining that
	\begin{equation*}
	\sum_{j = 1}^{n} \abs{\dotp{u_j}{v_k}} = \sqrt{2}\norm{a}_1 \leq \sqrt{2R} + 60\sqrt{2} n^2 \delta s_{\max} \sum_{r = R}^{n} r^{-5/2}
	\end{equation*}
	for any $R \geq 2 \vee k\sqrt{\delta} \vee 2^{5/4} 100^{1/4} n \sqrt{\delta s_{\max}}$. Bounding sum by integral, we have that if $R \geq 2$ then
	\begin{equation*}
	\sum_{r = R}^{n} r^{-5/2} \leq \int_{R - 1}^{n - 1} x^{-5/2} \,dx \leq \frac{2^{3/2}2}{3} R^{-3/2}.
	\end{equation*}
	Plugging this into the previous expression, we arrive at
	\begin{equation*}
	\sum_{j = 1}^{n} \abs{\dotp{u_j}{v_k}} \leq \sqrt{R} + 60 \sqrt{2} n^2 \delta s_{\max} \frac{2^{3/2}2}{3R^{3/2}}
	\end{equation*}
	for any $R \geq 2 \vee k\sqrt{\delta} \vee 2^{5/4} 100^{1/4} n \sqrt{\delta s_{\max}}$. The desired result then follows from~\eqref{eqn:incoherence_proof_1}.
\end{itemize}

% \paragraph{Davis-Kahan sin-theta theorem.}
%For a given $k \in [n]$ and $R \leq n$, decompose
%\begin{equation*}
%L = s_k v_k v_k^T + \sum_{k' \neq k}^{n} s_{k'} v_{k'} v_{k'}^T, ~~\textrm{and}~~ \widetilde{L} = \sum_{\ell:\abs{k - \ell} \leq R} \lambda_\ell u_\ell v_\ell^T + \sum_{\ell:\abs{k - \ell} > R} \lambda_\ell u_\ell u_\ell^T,
%\end{equation*}
%and denote $U_{k,R}^{\perp} = \sum_{\ell:\abs{k - \ell} > R} u_\ell u_\ell^T$. One consequence of the Davis-Kahan Theorem is the following: suppose for all $\ell: \abs{k - \ell} \geq R$, the eigenvalue $\lambda_{\ell}$ is excluded from the interval $(s_k - \delta,s_k + \delta)$ for some $\delta > 0$. Then 
%\begin{equation*}
%\sum_{\ell:\abs{k - \ell} > R} \dotp{v_k}{u_{\ell}}^2 \leq \frac{\norm{(U_{k,R}^{\perp})^T H v_k}_F^2}{\delta^2}
%\end{equation*}

\end{document}

### 2019-11-23 notes/weekly_notes/11.8.19.NotesForWeek.tex
added weekly notes

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\domain}{\mathcal{X}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\Partial}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 11/1/19 - 11/8/19}
\author{Alden Green}
\date{\today}
\maketitle

Let $P$ and $Q$ be distributions on $\Rd$, absolutely continuous with respect to Lebesgue measure with densities $p$ and $q$, and let $\mu = (P + Q)/2$. For fixed points $Z = \set{z_1,\ldots,z_N} \subset \mathcal{Z}$ for some subset $\mathcal{Z}$ of the support of $\mu$, suppose we observe $\ell = (\ell_1,\ldots,\ell_N) \in \Reals^n$, where $\ell_1,\ldots,\ell_N$ are independent Rademacher random variables with means
\begin{equation*}
\mathbb{E}[\ell_i] = \frac{p(z_i) - q(z_i)}{p(z_i) + q(z_i)}.
\end{equation*}

Let $\theta_p = (p(z_1),\ldots,p(z_N))$ and similarly $\theta_q = (q(z_1),\ldots,q(z_N))$; let $\theta_{\mu} = \frac{1}{2}\theta_p + \frac{1}{2}\theta_q$. Our goal: use the data $\ell,Z$ to test the following hypothesis:
\begin{equation*}
\textbf{H}_0: \theta_p = \theta_q ~~\textrm{versus}~~ \textbf{H}_a: d(\theta_p,\theta_q) > \epsilon 
\end{equation*}
for some metric $d$ on the space $\Reals^n$, and some $\epsilon > 0$. For a given test $\phi: \{-1,1\}^N \times \mathcal{Z} \to \{0,1\}$, we will evaluate our test error using the sum of Type I and Type II error
\begin{equation*}
\mathcal{R}(\phi;\theta_p,\theta_q) = \mathbb{E}_{\theta_\mu,\theta_\mu}(\phi) + \mathbb{E}_{\theta_p,\theta_q}(1 - \phi),
\end{equation*}
and the worst case risk
\begin{equation*}
\mathcal{R}(\phi,\epsilon,\Theta) = \sup_{\theta_p,\theta_q \in \Theta} \mathcal{R}(\phi; \theta_p, \theta_q)
\end{equation*}
where $\Theta$ is a subset of $\Reals^n$ and the supremum is over all vectors $\theta_p,\theta_q$ such that $d(\theta_p,\theta_q) \geq \epsilon$.

In particular, we propose and analyze the following projection-based test. Letting $G_{N,r}$ be the $r$-neighborhood graph over $Z$ with combinatorial Laplacian $L$, we write the spectral decomposition of $L$ as $L = V S V^T$, where $V = (v_1,\ldots,v_N)$ is $N \times N$ orthonormal matrix, and $S$ is a diagonal matrix with entries $s_1 \leq s_2 \leq \ldots \leq s_N$. Our test statistic $T$ depends on a tuning parameter $C > 0$ in the following manner:
\begin{equation*}
T = T(\ell,Z) = \frac{1}{N}\sum_{k: s_k \leq C} \left(\dotp{\ell}{v_k}\right)^2
\end{equation*}
and our test $\phi := \phi(\ell,Z)$ will simply be $\phi = \1\{T \geq \frac{\kappa}{N} + b\tau\}$ for a threshold $\tau > 0$ to be specified later.

\subsection{Moments of $T$.}
To upper bound the test error we must show that the fluctuations of $T$ (under null or alternative)  are small compared to the difference in means $\Ebb_{\theta_{p},\theta_q}(T) - \Ebb_{\theta_{\mu},\theta_{\mu}}(T)$. We therefore must compute the first and second moments of $T$, under null and alternative. To compute the moments of $T$, we decompose the vector $\ell$ into a mean term and a noise term. Letting $\Delta_{P,Q} = \frac{\theta_P - \theta_q}{2\theta_{\mu}}$, we have $\ell = \Delta_{P,Q} + \varepsilon$ where $\varepsilon = (\varepsilon_1,\ldots,\varepsilon_n)$ consists of $n$ independent mean-zero random variables, distributed
\begin{equation*}
\varepsilon_i = 
\begin{cases}
\frac{\theta_{Q,i}}{\theta_{\mu,i}}, ~ & \textrm{with probability $\frac{\theta_{P,i}}{\theta_{\mu,i}}$} \\
-\frac{\theta_{P,i}}{\theta_{\mu,i}}, ~ & \textrm{with probability $\frac{\theta_{Q,i}}{\theta_{\mu,i}}$}.
\end{cases}
\end{equation*}

\subsubsection{Mean of $T$.}
Let $\kappa$ be the largest integer $k \in [N]$ such that $s_k \leq C$. (Here $\kappa$ is a fixed quantity.) Then,
\begin{align*}
\Ebb(T) & = \frac{1}{N}\sum_{k = 1}^{\kappa} \Ebb\left[\dotp{v_k}{\ell}^2\right] \\
& = \frac{1}{N}\sum_{k = 1}^{\kappa} \Ebb\left[\dotp{\Delta_{P,Q} + \epsilon}{v_k}^2\right] \\
& = \frac{1}{N}\sum_{k = 1}^{\kappa} \bigl\{\dotp{\Delta_{P,Q}}{v_k}^2 + \Ebb\left[\dotp{\epsilon}{v_k}^2\right]\bigr\}
\end{align*}
For a given $k \in [\kappa]$, we have
\begin{align*}
\Ebb\left[\dotp{\epsilon}{v_k}_2^2\right] & = \sum_{i = 1}^{N} \sum_{j = 1}^N \Ebb\left[\varepsilon_i \varepsilon_j v_{k,i} v_{k,j}\right] \\
& = \sum_{i = 1}^{N} v_{k,i}^2 \Ebb\left[\varepsilon_i^2 \right]
\end{align*}
and by a direct computation of $\Ebb\left[\varepsilon_i^2 \right]$ we obtain
\begin{align*}
\mathbb{E}(T) & = \frac{1}{N}\left(\sum_{k = 1}^{\kappa} \dotp{\Delta_{P,Q}}{v_k}^2 + \sum_{k = 1}^{\kappa} \sum_{i = 1}^{N} v_{k,i}^2 \frac{\theta_{P,i} \theta_{Q,i}}{\theta_{\mu,i}^2}\right) \\
& = \frac{1}{N}\left(\sum_{k = 1}^{\kappa} \dotp{\Delta_{P,Q}}{v_k}^2 + \kappa - \sum_{k = 1}^{\kappa} \sum_{i = 1}^{N} v_{k,i}^2 \Delta_{P,Q,i}^2\right) \\
& = \frac{1}{N}\left(\sum_{k = 1}^{\kappa} \dotp{\Delta_{P,Q}}{v_k}^2 + \kappa - \sum_{i = 1}^{N} \Pi_{\kappa,ii} \Delta_{P,Q,i}^2\right) \\
& \geq \frac{1}{N}\left(\sum_{k = 1}^{\kappa} \dotp{\Delta_{P,Q}}{v_k}^2 + \kappa - \Pi_{\max} \norm{\Delta_{P,Q}}^2\right)
\end{align*}
Under the null, $\Delta_{P,Q} = 0$ and therefore
\begin{equation}
\label{eqn:mean_null}
\mathbb{E}_{\theta_{\mu},\theta_{\mu}}(T) = \frac{\kappa}{N}.
\end{equation}
Under the alternative, we have the following lower bound on the expectation of 
\begin{align*}
\sum_{k = 1}^{\kappa} \dotp{\Delta_{P,Q}}{v_k}^2 & = \sum_{k = 1}^{N} \dotp{\Delta_{P,Q}}{v_k}^2  - \sum_{k = \kappa + 1}^{N} \dotp{\Delta_{P,Q}}{v_k}^2 \\
& = \norm{\Delta_{P,Q}}_2^2 - \sum_{k = \kappa + 1}^{N} \frac{1}{s_k}\dotp{\Delta_{P,Q}}{v_k}^2 s_k \\
& \geq \norm{\Delta_{P,Q}}_2^2 - \frac{1}{C} \dotp{L\Delta_{P,Q}}{\Delta_{P,Q}}.
\end{align*}
leading to the lower bound
\begin{equation}
\label{eqn:mean_alternative}
\mathbb{E}_{\theta_p,\theta_q}(T) \geq \frac{\kappa}{N} + \frac{(1 - \Pi_{\max})}{N}\norm{\Delta_{P,Q}}^2 - \frac{1}{CN} \dotp{L\Delta_{P,Q}}{\Delta_{P,Q}}.
\end{equation}

\subsubsection{Variance of $T$.}
Let $V_{\kappa} = (v_1\ldots v_{\kappa})$ be the $n \times \kappa$ matrix containing the first $\kappa$ eigenvectors of $L$, and let $\Pi_{\kappa} = V_{\kappa} V_{\kappa}^T$. Observing that $T = \frac{1}{N}\dotp{\Pi_{\kappa}(\Delta_{P,Q} + \varepsilon)}{\Delta_{P,Q} + \varepsilon}$, we have
\begin{align}
\Var(T) & = \frac{1}{N^2}\Var(2\dotp{\Pi_{\kappa}\Delta_{P,Q}}{\varepsilon} + \dotp{\Pi_{\kappa}\varepsilon}{\varepsilon}) \nonumber \\
& = \frac{1}{N^2}\Bigl\{4\underbrace{\Var(\dotp{\Pi_{\kappa}\Delta_{P,Q}}{\varepsilon})}_{=: V_1} + 2\underbrace{\Cov(\dotp{\Pi_{\kappa}\Delta_{P,Q}}{\varepsilon},\dotp{\Pi_{\kappa}\varepsilon}{\varepsilon})}_{=: K_1} + \underbrace{\Var(\dotp{\Pi_{\kappa}\varepsilon}{\varepsilon})}_{=:V_2}\Bigr\}, \label{eqn:var}
\end{align}
and we now upper bound each of the three terms on the right hand side of the previous display.
\paragraph{Upper bound on $V_1$:}
Let $\Sigma := \Cov(\varepsilon)$ be the covariance matrix of $\varepsilon$. Noting that $\Sigma \preceq I$, we have
\begin{equation}
\label{eqn:var1}
\Var(\dotp{\Pi_{\kappa}\Delta_{P,Q}}{\varepsilon}) =\Delta_{P,Q}^T \Pi_{\kappa}^T \Sigma \Pi_{\kappa} \Delta_{P,Q} \leq \norm{\Pi_{\kappa}\Delta_{P,Q}}^2.
\end{equation}

\paragraph{Upper bound on $K_1$:}
Noting that $\Ebb(\dotp{\Pi_{\kappa} \Delta_{P,Q}}{\varepsilon}) = 0$, we have that
\begin{align*}
K_1 & = \Ebb\left[\dotp{\Pi_{\kappa} \Delta_{P,Q}}{\varepsilon} \dotp{\Pi_{\kappa} \varepsilon}{\varepsilon}\right] \\
& = \sum_{i = 1}^{N} \sum_{j = 1}^{N} \sum_{i' = 1}^{N} \Ebb\left[\varepsilon_i \varepsilon_j \Pi_{\kappa,ij} \varepsilon_{i'} (\Pi_{\kappa}\Delta_{P,Q})_i\right] \\
& = \sum_{i = 1}^{N} \Ebb\left[\varepsilon_i^3\right] \Pi_{\kappa,ii} (\Pi_{\kappa} \Delta_{P,Q})_i.
\end{align*}
Standard computations yield $\Ebb\left[\varepsilon_i^3\right]  = \frac{1}{2}(1 - \Delta_{P,Q,i}^2)(\Delta_{Q,P,i})$, and plugging in to the previous expression by Cauchy-Schwarz we have
\begin{align}
K_1 = \sum_{i = 1}^{N} \Ebb\left[\varepsilon_i^3\right] \Pi_{\kappa,ii}  (\Pi_{\kappa}\Delta_{P,Q})_i & = \frac{1}{2}\sum_{i = 1}^{N} (1 - \Delta_{P,Q,i}^2)(\Delta_{Q,P,i}) \Pi_{\kappa,ii}  (\Pi_{\kappa}\Delta_{P,Q})_i \nonumber \\
& \leq \norm{\Pi_{\kappa} \Delta_{P,Q}} \cdot \left(\sum_{i = 1}^{N} \Delta_{P,Q,i}^2 \Pi_{\kappa,ii}^2\right)^{1/2} \nonumber \\
& \leq \Pi_{\max} \cdot \norm{\Pi_{\kappa} \Delta_{P,Q}} \cdot \norm{\Delta_{P,Q}}  \label{eqn:var2} 
\end{align}

\paragraph{Upper bound on $V_2$:}
$V_2$ is a variance of a sum, which we re-express as the sum of covariances:
\begin{align*}
V_2 & = \Var(\dotp{\Pi_{\kappa} \varepsilon}{\varepsilon}) \\
& = \sum_{i,j,i',j' = 1}^{N} (\Pi_{\kappa})_{ij} (\Pi_{\kappa})_{i'j'} \Cov(\varepsilon_i \varepsilon_j, \varepsilon_{i'} \varepsilon_{j'}).
\end{align*}
This covariance will be non-zero only when $i = i' \neq j = j'$, $i = j' \neq j = i'$, or $i = i' = j = j'$, and therefore
\begin{align}
V_2 & = 2\sum_{i = 1}^{N}\sum_{j = 1}^{N} (\Pi_{\kappa})_{ij}^2 \Var(\varepsilon_i\varepsilon_j) + \sum_{i = 1}^{N} (\Pi_{\kappa})_{ii}^2 \Var(\varepsilon_i^2) \nonumber\\
& \leq 2\sum_{i = 1}^{N}\sum_{j = 1}^{N} (\Pi_{\kappa})_{ij}^2 + \sum_{i = 1}^{N} (\Pi_{\kappa})_{ii}^2 \leq 3 \mathrm{tr}(\Pi_{\kappa}^2) = 3\kappa. \label{eqn:var3}
\end{align}

\paragraph{Putting the pieces together.}
Plugging \eqref{eqn:var1}, \eqref{eqn:var2}, and \eqref{eqn:var3} back into \eqref{eqn:var}, we obtain
\begin{equation}
\label{eqn:var_5}
\Var(T) \leq \frac{1}{N^2}\left\{4\norm{\Pi_{\kappa} \Delta_{P,Q}}^2 + \Pi_{\max} \cdot \norm{\Pi_{\kappa} \Delta_{P,Q}} \cdot \norm{\Delta_{P,Q}} + 3\kappa \right\}.
\end{equation}

\subsection{Type I and Type II error.}

We translate our bounds on the mean and variance of $T$ to bounds on the Type I and Type II error of our test $\phi$ using Chebyshev's inequality. Recall that our test statistic is of the form $\phi = \1\{T \geq \frac{\kappa}{N} + b\tau\}$, and let $\tau = \frac{\sqrt{3\kappa}}{N}$.

\paragraph{Type I error.}
We apply Chebyshev's inequality, and obtain
\begin{align*}
\Pbb_{\theta_{\mu},\theta_{\mu}}\left(T \geq \frac{\kappa}{N} + b\tau\right) & = \Pbb_{\theta_{\mu},\theta_{\mu}}\left(T - \frac{\kappa}{N} \geq b\tau - \Ebb(T)\right) \\
& = \Pbb_{\theta_{\mu},\theta_{\mu}}\left( (T - \frac{\kappa}{N})^2 \geq b^2\tau^2 \right) \\
& \leq \frac{\Var_{\theta_{\mu},\theta_{\mu}}(T)}{b^2\tau^2} = \frac{1}{b^2}.
\end{align*}

\paragraph{Type II error.}
To obtain meaningful bounds on Type I error, we must assume some separation between $\theta_P$ and $\theta_Q$. In particular, letting $d(\theta_P,\theta_Q) := \norm{\Delta_{P,Q}}^2$ we assume
\begin{equation}
\label{eqn:dist_thetaP_thetaQ}
\frac{1}{N}d(\theta_P,\theta_Q) \geq \frac{1}{(1 - \Pi_{\max})} \left(\frac{1}{CN}\norm{D \Delta_{P,Q}}^2 + 2b\tau\right).
\end{equation}
Note that this implies
\begin{equation*}
\mathbb{E}_{\theta_P,\theta_Q}(T) - \left(\frac{\kappa}{N} + b\tau\right) \geq \frac{(1 - \Pi_{\max})}{N} \norm{\Delta_{P,Q}}^2 - \frac{1}{CN}\norm{D \Delta_{P,Q}}^2 - b\tau \geq b\tau \vee \frac{(1 - \Pi_{\max})}{2N} \norm{\Delta_{P,Q}}^2.
\end{equation*}
Now, we apply Chebyshev's inequality:
\begin{align*}
\Pbb\left(T \leq \frac{\kappa}{N} + b\tau\right) & = \Pbb\bigl(T - \Ebb[T] \leq \frac{\kappa}{N} + b\tau - \Ebb[T]\bigr) \\
& = \Pbb\bigl( (T - \Ebb[T])^2 \leq (\Ebb[T] - \frac{\kappa}{N} - b\tau)^2 \bigr) \\
& \leq \frac{\Var(T)}{(\Ebb[T] - \frac{\kappa}{N} - b\tau)^2} \\
& \leq \frac{1}{N^2}\frac{4\norm{\Pi_{\kappa} \Delta_{P,Q}}^2 + \Pi_{\max} \cdot \norm{\Pi_{\kappa} \Delta_{P,Q}} \cdot \norm{\Delta_{P,Q}} + 3\kappa}{b^2\tau^2 \vee \frac{(1 - \Pi_{\max})^2}{4N^2} \norm{\Delta_{P,Q}}^4} \\
& \leq \frac{16}{(1 - \Pi_{\max})^2 \norm{\Delta_{P,Q}}^2} \vee \frac{4\Pi_{\max}}{(1 - \Pi_{\max})^2 \norm{\Delta_{P,Q}}^2} \vee \frac{1}{b^2} \\
& \leq \frac{3}{b^2}
\end{align*}
where the last line follows by \eqref{eqn:dist_thetaP_thetaQ} whenever $N \geq \sqrt{48}$.

\section{Upper bounds on $\Pi_{\max}$.}
Clearly, the critical radius established by inequality~\eqref{eqn:dist_thetaP_thetaQ} is only meaningful if $1 - \Pi_{\max} > 0$. In fact, we will want to show that $\Pi_{\max} < \frac{1}{2}$ so the factor of $(1 - \Pi_{\max})^{-1}$ does not affect the rate at which the critical radius shrinks. We review what $\Pi_{\max}$ looks like on some common graphs.

\subsection{Chain.}
Suppose $G := G_{\mathrm{1N}}$ is a length-$N$ chain, i.e. a 1d grid graph on vertices $V = 1:N$. The eigenvectors of $G$ are given by the discrete Fourier transform;
\begin{equation*}
v_{1,i} = \frac{1}{\sqrt{N}},~~ v_{k,i} =\sqrt{\frac{2}{N}} \cos\left(\frac{(i - .5)(k - 1)\pi}{N}\right)~ \textrm{for $i = 1,\ldots,N$, k = $2,\ldots,N$.}
\end{equation*}
Clearly $\Pi_{\max} \leq \frac{2\kappa}{N}$, so that whenever $\kappa < \frac{N}{4}$ we have $\Pi_{\max} < \frac{1}{2}$ as desired.

\subsection{Grid.}
Now suppose $G := G_{\mathrm{dN}}$ is a $d$-dimensional grid graph on $N$ nodes, i.e. $G = G_{1M} \otimes \cdots \otimes G_{1M}$ for $M = \frac{N}{d}$ (assume $M$ is an integer for simplicity). The product structure of $G$ yields an expression for its eigenvectors from the eigenvectors of $G_{\mathrm{1M}}$. In particular, for a given $i = (i_1,\ldots,i_d) \in [M]^d$ and $k = (k_1,\ldots,k_d) \in [M]^d$, we have
\begin{equation*}
v_{k,i} = \prod_{j = 1}^{d} \sqrt{\frac{2}{M}} \cos\left(\frac{(i_j - .5)(k_j - 1)\pi}{M}\right)
\end{equation*}
and therefore $v_{k,i}^2 \leq \frac{2}{M^{d}} = \frac{2}{N}$ for all $i,k \in [M]^d$. So we arrive at the same bound $\Pi_{\max} \leq \frac{2\kappa}{N}$ and again $\kappa < \frac{N}{4}$ implies $\Pi_{\max} < \frac{1}{2}$ as desired.

\subsection{Spectral similarity.}
Suppose we have graphs $G$ and $H$ with Laplacians $L_G$ and $L_H$, such that $G$ and $H$ are $\epsilon$ spectrally-similar, meaning
\begin{equation}
\label{eqn:spectral_similarity}
(1 - \epsilon) x^T L_H x \leq x^T L_G x \leq (1 + \epsilon) x^T L_H x,~~ \textrm{for all $x \in \Reals^N$.}
\end{equation}
Write $L_G = V \Lambda V^T$ and $L_H = U S U^T$ for the spectral decompositions of $G$ and $H$, respectively, and assume that we have
\begin{equation*}
\max_{(i,l) \in [N]^2} u_{l,i}^2 \leq \frac{2}{N},
\end{equation*}
which we have already shown holds for $H = G_{dN}$. For each $k$, and for any $L(k), R(k) \in [N]$, write
\begin{equation*}
v_{k} = \sum_{l = R(k)}^{L(k)}\alpha_{k,l} u_l +  \textrm{proj}_{U_{L(k)}^{\perp}}(v_k);
\end{equation*}
as a result, we have that
\begin{align*}
v_{k,i}^2 & \leq 2\sum_{h = R(k)}^{L(k)} \sum_{l = R(k)}^{L(k)}\alpha_{k,h}\alpha_{k,l} u_{h,i} u_{l,i} + 2(\textrm{proj}_{U_{R(k):L(k)}^{\perp}}(v_k))_i^2 \\
& \leq \frac{4}{N} \norm{\alpha}_1^2 + 2\norm{ \textrm{proj}_{U_{L(k)}^{\perp}}(v_k)}^2 \\
& \leq \frac{4}{N} (L(k) - R(k)) + 2\norm{ \textrm{proj}_{U_{R(k):L(k)}^{\perp}}(v_k)}^2
\end{align*}
where the last inequality follows since $\norm{\alpha}_2 \leq 1$, and the inequality $\norm{\alpha}_1 \leq \sqrt{L - R} \norm{\alpha}_2$, since $\alpha$ is a length $L - R$-vector.

\textcolor{red}{TODO:} Upper bound $\norm{ \textrm{proj}_{U_{L(k)}^{\perp}}(v_k)}^2$ using~\eqref{eqn:spectral_similarity}, and choose $L(k),R(k)$ to make the sum in the previous display as small as possible.
\section{Notation.}

We write $\Pi_{\max} = \max_{i = 1,\ldots,n} \Pi_{\kappa,ii}$ where we recall $\Pi_{\kappa} = V_{\kappa} V_{\kappa}^T$ and therefore $\Pi_{\kappa,ii} = \sum_{k = 1}^{\kappa} v_{k,i}^2$.

We write $D$ for the incidence matrix of $L$.

\end{document}

### 2019-12-10 notes/graph_testing.tex
added high-level summary and project-level notes

--- 2019-02-08 notes/graph_testing.tex
@@
 \newcommand{\Identity}{\mathbb{I}}
-\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
+\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
 \newcommand{\convprob}{\overset{p}{\to}}
@@
 \newcommand{\myexp}[1]{\exp \{ #1 \}}
-\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
-\newcommand{\dotp}[2]{\langle #1 , #2 \rangle}
 \newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
 \newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
-\newcommand{\defeq}{\overset{\mathrm{def}}{=}}
-\newcommand{\convweak}{\overset{w}{\rightharpoonup}}
-\newcommand{\dive}{\mathrm{div}}
-\newcommand{\Bin}{\mathrm{Bin}}
+\newcommand{\ext}[1]{\widetilde{#1}}
+\newcommand{\set}[1]{\left\{#1\right\}}
+\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
+\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
+\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
+\newcommand{\Var}{\mathrm{Var}}
+\newcommand{\Cov}{\mathrm{Cov}}
+\newcommand{\Xsetiam}{\mathrm{diam}}
 
@@
 \newcommand{\vol}{\text{vol}}
-\newcommand{\Bal}{\textrm{Bal}}
-\newcommand{\Cut}{\textrm{Cut}}
-\newcommand{\Ind}{\textrm{Ind}}
-\newcommand{\set}[1]{\left\{#1\right\}}
-\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
-\newcommand{\Perp}{\perp \! \! \! \perp}
-\newcommand{\Naturals}{\mathbb{N}}
+\newcommand{\spansp}{\mathrm{span}~}
+\newcommand{\1}{\mathbf{1}}
 
-\newcommand{\Linv}{L^{\dagger}}
-\newcommand{\tr}{\text{tr}}
-\newcommand{\h}{\textbf{h}}
-% \newcommand{\l}{\ell}
-\newcommand{\x}{\textbf{x}}
-\newcommand{\y}{\textbf{y}}
-\newcommand{\bl}{\bm{\ell}}
-\newcommand{\bnu}{\bm{\nu}}
-\newcommand{\Lx}{\mathcal{L}_X}
-\newcommand{\Ly}{\mathcal{L}_Y}
+\newcommand{\Linv}{L^{\Xsetagger}}
 \DeclareMathOperator*{\argmin}{argmin}
+\DeclareMathOperator*{\argmax}{argmax}
 
+\newcommand{\emF}{\mathbb{F}_n}
 \newcommand{\emG}{\mathbb{G}_n}
-\newcommand{\A}{\mathcal{A}}
+\newcommand{\emP}{\mathbb{P}_n}
 \newcommand{\F}{\mathcal{F}}
-\newcommand{\G}{\mathcal{G}}
-\newcommand{\X}{\mathcal{X}}
+\newcommand{\D}{\mathcal{D}}
+\newcommand{\R}{\mathcal{R}}
 \newcommand{\Rd}{\Reals^d}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\E}{\mathcal{E}}
 
-%%% Matrix related notation
-\newcommand{\Xbf}{\mathbf{X}}
-\newcommand{\Ybf}{\mathbf{Y}}
-\newcommand{\Zbf}{\mathbf{Z}}
-\newcommand{\Abf}{\mathbf{A}}
-\newcommand{\Dbf}{\mathbf{D}}
-\newcommand{\Wbf}{\mathbf{W}}
-\newcommand{\Lbf}{\mathbf{L}}
-\newcommand{\Ibf}{\mathbf{I}}
-\newcommand{\Bbf}{\mathbf{B}}
+%%% Vectors
+\newcommand{\thetast}{\theta^{\star}}
 
-%%% Vector related notation
-\newcommand{\lbf}{\bm{\ell}}
-\newcommand{\fbf}{\mathbf{f}}
+%%% Matrices
+\newcommand{\X}{X} % no bold
+\newcommand{\Y}{Y} % no bold
+\newcommand{\Z}{Z} % no bold
+\newcommand{\Lgrid}{L_{\grid}}
+\newcommand{\Xsetgrid}{D_{\grid}}
+\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}
 
-%%% Set related notation
-\newcommand{\Dset}{\mathcal{D}}
-\newcommand{\Aset}{\mathcal{A}}
-\newcommand{\Wset}{\mathcal{W}}
+%%% Sets and classes
+\newcommand{\Xset}{\mathcal{X}}
+\newcommand{\Sset}{\mathcal{S}}
+\newcommand{\Hclass}{\mathcal{H}}
+\newcommand{\Pclass}{\mathcal{P}}
+\newcommand{\Leb}{\mathcal{L}}
 
-%%% Distribution related notation
+%%% Distributions and related quantities
 \newcommand{\Pbb}{\mathbb{P}}
+\newcommand{\Ebb}{\mathbb{E}}
 \newcommand{\Qbb}{\mathbb{Q}}
-% \newcommand{\Pr}{\mathrm{Pr}}}
 
-%%% Functionals
-\newcommand{\1}{\mathbf{1}}
+%%% Operators
+\newcommand{\Tadj}{T^{\star}}
+\newcommand{\Xsetive}{\mathrm{div}}
+\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
+\newcommand{\gradient}{\mathcal{D}}
+\newcommand{\Hessian}{\mathcal{D}^2}
+\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
+
+%%% Misc
+\newcommand{\grid}{\mathrm{grid}}
+\newcommand{\critr}{R_n}
+\newcommand{\Xsetx}{\,dx}
+\newcommand{\Xsety}{\,dy}
+\newcommand{\Xsetr}{\,dr}
+\newcommand{\Xsetxpr}{\,dx'}
+\newcommand{\Xsetypr}{\,dy'}
+\newcommand{\wt}[1]{\widetilde{#1}}
+\newcommand{\spec}{\mathrm{spec}}
+
+%%% Order of magnitude
+\newcommand{\soom}{\sim}
+
+% \newcommand{\span}{\textrm{span}}
 
@@
 
-\paragraph{Two-sample non-parametric hypothesis-testing problem.}
+This document details the current status of the graph testing project.
+We divide by section based on the testing problem under consideration.
+At the end of each section, we list some areas we are interested in investigating.
+All proofs are left until the end.
 
-For fixed integers $n_1 + n_2 = n$, let $\Xbf = \set{x_1, \ldots, x_{n_1}} \subset \Rd$ and $\Ybf = \set{y_1, \ldots, y_{n_2}}$ be sampled i.i.d from distributions $\Pbb$ and $\Qbb$ with density functions $p$ and $q$, respectively, both with support on $D \subset \Rd$.
-Our statistical problem is testing the null hypothesis $H_0: \Pbb = \Qbb$ vs.
-the alternative $H_1: \Pbb \neq \Qbb$, where our knowledge of $\Pbb$ and $\Qbb$ come from the samples $\Xbf$ and $\Ybf$.
+Throughout, let $X = \{x_1,
+\ldots, x_n\}$ be a sample drawn i.i.d. from a distribution $P$ on $\Rd$,
+with density~$p$.  For a radius $r > 0$, we define $G_{n,r}=(V,E)$ to be the
+\emph{$r$-neighborhood graph} of $X$, an unweighted, undirected graph with
+vertices $V=X$, and an edge $(x_i,x_j) \in E$ if and only if $K_r(x_i,x_j) = \norm{x_i -x_j} \leq r$, where $\norm{\cdot}$ is the Euclidean norm. We denote by $A \in
+\Reals^{n \times n}$ the adjacency matrix, with entries $A_{uv} = 1$ if
+$(u,v) \in E$ and $0$ otherwise.  We also denote by $D$ the diagonal degree
+matrix, with entries $D_{uu} := \sum_{v \in V} A_{uv}$. The graph Laplacian is $L = D - A$, and we write its spectral decomposition as $L = V S V^T$.
 
-\paragraph{Integral Probability Metric.}
-Given $\F$ a class of real-valued bounded measurable functions on $D$, the \emph{integral probability metric} between $\Pbb$ and $\Qbb$ with respect to $\F$ is
+\section{Regression goodness-of-fit testing with random design.}
+
+Let $P$ be a distribution with density $p$ supported on $\mathcal{X} \subseteq \Reals^d$. Suppose we observe random design points $X = \set{x_1,\ldots,x_n} \sim P$, and additionally responses
+\begin{equation}
+\label{eqn:regression_known_variance}
+y_i = f(x_i) + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
+\end{equation}
+
+We wish to distinguish
 \begin{equation*}
-\gamma_{\F}(\Pbb,\Qbb) = \sup_{f \in \F}\abs{\int_{D} f d\Pbb - \int_{D} f d\Qbb}
+\mathbf{H}_0: f = f_0 := 0 \quad \textrm{vs} \quad \mathbf{H}_{\textrm{a}}: f \neq f_0
+\end{equation*}
+We will evaluate our performance using worst-case risk: for a given function class $\mathcal{H}$ and test function $\phi: \Reals^n \to \set{0,1}$, let
+\begin{equation*}
+\mathcal{R}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, f \neq f_0} \Ebb_f(1 - \phi).
+\end{equation*}
+The worst-case risk may be quite close to $1$ unless we require some separation between null and alternative spaces. A more realistic measure of performance is therefore
+\begin{equation*}
+\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, \norm{f - f_0}_2 \geq \epsilon} \Ebb_f(1 - \phi).
 \end{equation*}
 
-One such IPM that has not received close attention until surprisingly recently is the \emph{(weighted) Sobolev IPM}.
-For $f \in L^2(\Dset)$ and $\rho$ a density function over $\Dset$, the ($\rho^2$-weighted) Sobolev 1,2 norm of $f$ is given by
+\subsection{Test Statistics.}
+We list the test statistics we use for the regression testing problem.
+
+\paragraph{Eigenvector projection test statistic.}
+Let $VSV^T$ be the spectral decomposition of the Laplacian matrix $L$ of the neighborhood graph $G_{n,r}$.
+To test whether $f = f_0$, we propose the following \emph{eigenvector projection} test statistic:
+\begin{equation}
+\label{eqn:graph_spectral_projections}
+T_{\mathrm{spec}} := \frac{1}{n} \sum_{k = 0}^{\kappa} \left(\sum_{i = 1} v_i y_i\right)^2
+\end{equation}
+
+\subsection{Current Results.}
+
+In Theorem~\ref{thm:sobolev_testing_rate} we show that under some typical regularity conditions on $P$, the test $\phi_{\textrm{spec}} := \1\{T_{\mathrm{spec}} \geq \tau\}$ is, up to log factors, a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};R)$.
+To conveniently state our results we introduce the notation
 \begin{equation*}
-\norm{f}_{1,2,\rho^2} := 
+h(a,d) = a(d+2) + (1 + 2/d),~~\textrm{for $a > 0$}
+\end{equation*}
+
+\begin{theorem}
+	\label{thm:sobolev_testing_rate}
+	Let $b \geq 1$ and $a > 0$ be fixed constants, and let $d < 4$.
+Suppose that $\Pbb$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p$ bounded above and below by constants, i.e
+	\begin{equation*}
+	0 < p_{\min} < p(x) < p_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
+	\end{equation*}
+	Then the following statement holds: if the test $\phi_{\spec}$ is performed with parameter choices 
+	\begin{equation*}
+	r = \log^a n \cdot \left(\frac{\log n}{n}\right)^{1/d}, ~\kappa = n^{2d/(4 + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	\end{equation*}
+	then there exists a constant $c$ which may depend on $d,R,p_{\max},b$ and $a$ but is independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
+	\begin{equation}
+	\label{eqn:sobolev_testing_rate}
+	\epsilon^2 \geq c \cdot b \cdot n^{-4/(4 + d)} (\log n)^{h(a,d)}
+	\end{equation}
+	the worst-case risk is upper bounded
+	\begin{equation}
+	\label{eqn:sobolev_testing_rate_1}
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; \mathcal{W}^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}.
+	\end{equation}
+\end{theorem}
+The testing rate~\eqref{eqn:sobolev_testing_rate} matches the minimax critical radius up to a factor of $\log^{h(a,d)}n$.
+
+When $d \geq 4$ the compact embedding
+\begin{equation*}
+W_d^{1,2}(\mathcal{X}) \subseteq \mathcal{L}_d^4(\mathcal{X}) 
+\end{equation*}
+does not hold (as it does when $d < 4$). However, if we directly assume $f \in \mathcal{L}_d^4(\mathcal{X};R)$ (regardless of $d$) we obtain the following result.
+\begin{proposition}
+	\label{prop:L4_testing_rate}
+	If $f \in \mathcal{L}_d^4(\mathcal{X};R)$, there exists a constant $c$ such that if
+	\begin{equation}
+	\label{eqn:L4_testing_rate}
+	\epsilon^2 > b \cdot n^{-1/2}
+	\end{equation}
+	then the test
+	\begin{equation*}
+	\phi_{\mathrm{mean}} = \1\{\frac{1}{n}\sum_{i = 1}^{n} y_i^2 \geq 1\}
+	\end{equation*}
+	has worst-case risk
+	\begin{equation*}
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}. 
+	\end{equation*}
+\end{proposition}
+
+Note that when $d < 4$ the rate~\eqref{eqn:sobolev_testing_rate} is sharper than \eqref{eqn:L4_testing_rate}.
+
+\subsection{Areas to Investigate}
+
+\begin{enumerate}[(i)]
+	\item When $d \geq 4$, is it true that there does not exist a uniformly consistent
+	test over the Sobolev ball $W_d^{1,2}(\mathcal{X};R)$?
+	\item Does the testing rate of $\phi_{\mathrm{spec}}$ over Sobolev spaces $W_d^{s,2}(\mathcal{X};R)$ match the minimax rate, for an appropriate choice of kernel $K$?
+	\item Assume the distribution $P$ is supported on a manifold $\mathcal{M}$ of intrinsic dimension $s < d$.
+Does $\mathrm{\phi_{\mathrm{spec}}}$ display adaptivity to the intrinsic dimension of $\mathcal{M}$?
+	\item Assume that $f$ belongs to the Holder space $C_d^s(\mathcal{X})$.
+Moreover, suppose that instead of observing ${y_i}$ according to the regression testing model \eqref{eqn:regression_known_variance}, we observe
+	\begin{equation*}
+	y_i = f(x_i) + \sigma \varepsilon_i, ~\varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
+	\end{equation*}
+	where $\sigma > 0$ is unknown.
+When $d \geq 4$, what are the minimax regression testing rates over $C_d^1(\mathcal{X};R)$?
+Is the test $\phi_{\mathrm{spec}}$ minimax optimal, when the tuning parameters $r$ and $\kappa$ are appropriately chosen?
+\end{enumerate}
+
+\section{Two-sample density testing.}
+In the two-sample density testing problem, we observe independent samples $Z = z_1,\ldots,z_N \sim P$ and $Y = y_1,\ldots,y_M \sim Q$, where $P$ and $Q$ are distributions over $\Reals^d$ with densities $p$ and $q$, respectively, and $N \sim \textrm{Bin}(n,1/2)$.
+Our goal is to distinguish the hypotheses
+\begin{equation*}
+\mathbf{H}_0: P = Q \quad \textrm{vs} \quad \mathbf{H}_{\textrm{a}}: P \neq Q
+\end{equation*}
+and we again evaluate our performance using worst-case risk; letting $\phi:\Reals^{N + M} \to \{0,1\}$, 
+\begin{equation*}
+\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \inf_{p \in \mathcal{H}}\Ebb_{p,p}(\phi) + \sup_{\substack{p,q \in \mathcal{H} \\ \norm{p - q}_{\Leb^2} \geq \epsilon}} \Ebb_{p,q}(1 - \phi).
+\end{equation*}
+
+\subsection{Test statistics.}
+
+We suggest several two-sample test statistics.
+
+\paragraph{Eigenvector projection test statistic.}
+
+It is straightforward to adapt the test statistic $T_{\mathrm{spec}}$ to the two-sample testing problem.
+Concatenate the samples $Z$ and $Y$ in $X = (z_1,\ldots,z_N,y_1,\ldots,y_M)$, and define $T_{\mathrm{spec}}^{(2)}$ to be
+\begin{equation}
+\label{eqn:graph_spectral_projections_2}
+T_{\mathrm{spec}}^{(2)} := \frac{1}{n} \sum_{k = 0}^{\kappa} \left(\sum_{i = 1}^{n} v_i a_i\right)^2, ~~\textrm{where}~~ a = (\underbrace{N^{-1},\ldots,{N^{-1}}}_{\textrm{length } N},\underbrace{-M^{-1},\ldots,-M^{-1}}_{\textrm{length } M})
+\end{equation}
+
+For convenience, we state our following two test statistics with respect to the empirical norm $\norm{\theta}_n = n^{-1/2}\norm{\theta}_2$ for $\theta \in \Reals^n$.
+They will each depend on a tuning parameter $\lambda > 0$.
+\paragraph{Graph Sobolev IPM.}
+Letting $C_n := nr^{(d + 2)/2}$ and
+\begin{equation*}
+\Theta_{1,2} := \{\theta \in \Reals^n:~ C_n^{-1} \norm{B\theta}_2 \leq 1\} 
+\end{equation*}
+we define the \emph{graph Sobolev IPM} to be
+\begin{equation}
+\label{eqn:sobolev_IPM}
+T_{\textrm{sob}} := \sup_{\substack{\theta \in \Theta_{1,2} \\ \lambda \norm{\theta}_n \leq 1}} \abs{\frac{1}{n}\sum_{i = 1}^{n} a_i \theta_i}. 
+\end{equation}
+
+\paragraph{Graph Total Variation IPM.}
+Letting $C_n' := n^{2}r^{(d + 1)}$ and 
+\begin{equation*}
+\Theta_{1,1} := \{\theta \in \Reals^n:~ (C_n')^{-1} \norm{B\theta}_1 \leq 1\}
+\end{equation*}
+we define the \emph{graph Total Variation} IPM to be
+\begin{equation}
+\label{eqn:total_variation_IPM}
+T_{\mathrm{TV}} := \sup_{\substack{\theta \in \Theta_{1,1}, \\ \lambda \norm{\theta}_n \leq 1} } \abs{\frac{1}{n}\sum_{i = 1}^{n} a_i \theta_i}, \quad
+\end{equation}
+
+\subsection{Current results.}
+
+\textcolor{red}{WARNING:} While I have proved all the parts I believe are required for the below theorem, I haven't actually put together the pieces yet.
+
+In Theorem~\ref{thm:twosample_sobolev_testing_rate} we show that under some typical regularity conditions on $P$, the test $\phi_{\textrm{spec}}^{(2)} := \1\{T_{\mathrm{spec}}^{(2)} \geq \tau\}$ is, up to log factors, a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};R)$ when $d = 1$.
+
+\begin{theorem}
+	\label{thm:twosample_sobolev_testing_rate}
+	Let $b \geq 1$ and $a > 0$ be fixed constants, and let $d = 1$.
+Suppose that $P$ and $Q$ are absolutely continuous probability measures over $\mathcal{X} = [0,1]$ with density functions $p$ and $q$ bounded above and below by constants, i.e
+	\begin{equation*}
+	0 < p_{\min},q_{\min} < p(x),q(x) < p_{\max},q_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
+	\end{equation*}
+	Then the following statement holds: if the test $\phi_{\spec}^{(2)}$ is performed with parameter choices 
+	\begin{equation*}
+	r = \log^a n \cdot \left(\frac{\log n}{n}\right), ~\kappa = n^{2/5}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	\end{equation*}
+	then there exists a constant $c$ which may depend on $R,p_{\max},q_{\max},b$ and $a$ but is independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
+	\begin{equation}
+	\label{eqn:twosample_sobolev_testing_rate}
+	\epsilon^2 \geq c \cdot b \cdot n^{-4/5} (\log n)^{h(a,1)}
+	\end{equation}
+	the worst-case risk is upper bounded
+	\begin{equation}
+	\label{eqn:twosample_sobolev_testing_rate_1}
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; \mathcal{W}^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}.
+	\end{equation}
+\end{theorem}
+
+We prove Theorem~\ref{thm:twosample_sobolev_testing_rate} by relating the density testing problem to a regression testing problem with a certain type of structured noise, and then proceeding along similar lines to the proof of Theorem~\ref{thm:sobolev_testing_rate}.
+To pursue this strategy, we require the eigenvectors to satisfy a certain type of incoherence condition; this is in constrast to the regression testing problem with known variance, where we did not require the eigenvectors to be smooth in any sense.
+
+\subsection{Areas to Investigate}
+
+\begin{enumerate}[(i)]
+	\item When $1 < d < 4$, is the test $\phi_{\spec}^{(2)}$ minimax optimal?
+	\item Are either or both of the test statistics \eqref{eqn:graph_spectral_projections_2}-\eqref{eqn:sobolev_IPM} minimax optimal over $C_d^1(\mathcal{X};R)$ for all values of $d$?
+	\item Are either or both of the test statistics \eqref{eqn:graph_spectral_projections_2}-\eqref{eqn:sobolev_IPM} minimax optimal over $W_d^{1,2}(\mathcal{X};R)$ when $d \leq 4$?
+	\item Is the test statistic \eqref{eqn:graph_spectral_projections_2}, computed over a graph with suitable choice of kernel $K$, minimax optimal over higher order derivative classes $C_d^s$ and $W^{s,2}$?
+	\item Modify the test statistic \eqref{eqn:sobolev_IPM} by replacing the function class $\Theta_{1,2}$ with
+	\begin{equation}
+	\Theta_{s,2} := \{\theta \in \Reals^n:~ C_n^{-1} \norm{B^{(s)}\theta}_2 \leq 1\} 
+	\end{equation}
+	Is the modified test statistic, computed over a graph with suitable choice of kernel $K$, minimax optimal over higher order derivative classes $C_d^s$ and $W^{s,2}$?
+	\item What is the minimax testing rate over $BV_d^{1}(\mathcal{X};R)$? Does it exhibit a phase transition analogous to the minimax estimation rate over bounded variation spaces?
+	\item Is the test statistic \eqref{eqn:total_variation_IPM} minimax optimal over $BV_d^{1}(\mathcal{X};R)$?
+	\item Is the test statistic \eqref{eqn:total_variation_IPM} minimax optimal over Sobolev and Holder function classes?
+\end{enumerate}
+
+\section{Definitions}
+
+Here we collect definitions of some common function spaces and graph operators.
+
+\subsection{Function Spaces}
+
+\paragraph{Lebesgue spaces.}
+
+We say a Borel measurable function $f: \mathcal{X} \to \Reals$ is in the space $\mathcal{L}^p(\mathcal{X})$ for $1 \leq p < \infty$ if 
+$$\norm{f}_{\mathcal{L}^p(\mathcal{X})} := \int_{\mathcal{X}} \abs{f(x)}^p \,dx < \infty$$
+and we let 
+\begin{equation*}
+\mathcal{L}^p(\mathcal{X};R) = \set{f \in \mathcal{L}^p(\mathcal{X}): \norm{f}_{\mathcal{L}^p(\mathcal{X})} < R}
+\end{equation*}
+be a ball in the Lebesgue space.
+
+\paragraph{Holder spaces.}
+
+For a given $s > 0$, the $s$th Holder norm is given by
+\begin{equation*}
+\norm{f}_{C_d^{s}(\mathcal{X})} := \sum_{\abs{\alpha} \leq s} \norm{D^{\alpha}f}_{\infty} + \sum_{\abs{\alpha} = s} \sup_{x,y \in \mathcal{X}} \frac{\abs{D^{\alpha}f(y) - D^{\alpha}f(x)}}{\norm{x - y}_2}
+\end{equation*}
+and the $s$th Holder space $C_d^{s}(\mathcal{X})$ consists of all functions which are $s$ times continuously differentiable with finite $s$ Holder norm.
+Denote the Holder  ball by $C_d^{s}(\mathcal{X},R) = \set{f \in C_d^{s}(\mathcal{X}): \norm{f}_{C_d^{s}(\mathcal{X})} \leq R}$.
+
+\paragraph{Sobolev spaces.}
+
+For a given $s > 0$, the Sobolev space $W_d^{s,2}(\mathcal{X})$ consists of all functions $f \in \mathcal{L}^2(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^2(\mathcal{X})$.
+The Sobolev $\{s,2\}$ norm is then
+\begin{equation*}
+\norm{f}_{W_d^{s,2}(\mathcal{X})}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^2 \,dx
+\end{equation*}
+and for a given $L > 0$, the corresponding ball is $W_d^{s,2}(\Xset; L) = \set{f: \norm{f}_{W^{s,2}(\Xset)} \leq L}$.
+
+\paragraph{Bounded Variation spaces.}
+
+For a function $f \in L^1(\mathcal{X})$ the \emph{total variation} semi-norm of $f$ is
+\begin{equation*}
+TV(f;\mathcal{X}) := \sup \left\{ \int_{\mathcal{X}} f \, \Xsetive \, \psi \,dx : \psi \in C_c^1(\mathcal{X}; \Reals^d), \abs{\psi} \leq 1 \right\};
+\end{equation*}
+and we write $BV_d(\mathcal{X})$ for the subset of functions $f \in L^1(\mathcal{X})$ which have bounded norm
+\begin{equation*}
+\norm{f}_{BV_d(\mathcal{X})} := \norm{f}_{\infty} + TV(f;\mathcal{X}).
+\end{equation*}
+For a given $R > 0$, the corresponding ball is $BV_d^{1}(\mathcal{X};R) = \set{f: \norm{f}_{BV_d(\mathcal{X})} \leq R}$.
+
+\subsection{Graph Operators.}
+Let $s \geq 1$ be an integer. The $s$th-order difference operator on $G_{n,r}$, denoted $B^{(s)}$, is defined by
+\begin{equation*}
+B^{(s)} :=
 \begin{cases}
-\int_{\Dset}\norm{\nabla_x f(x)}^2 \rho^2 dx,~ & \text{$f \in H^1(\Dset)$} \\
-\infty,~ & \text{$f \in L^2(\Dset) \setminus H^1(\Dset)$}
+L^{s/2},& ~~ s \textrm{ even} \\
+BL^{(s - 1)/2},& ~~ s \textrm{ odd.}
 \end{cases}
 \end{equation*}
-where $H^1(\Dset)$ is \textcolor{red}{the Sobolev space of $L^2(\Dset)$ functions with weak derivative $\nabla_x f(x) \in L^2(\Dset)$.}
 
-Consider the unit ball of $\norm{\cdot}_{1,2,\rho^2}$,
-\begin{equation*}
-\Wset^{1,2}(\Dset,\rho^2) := \set{f: \norm{f}_{1,2,\rho^2} \leq 1}.
-\end{equation*}
-The weighted Sobolev IPM is simply $\gamma_{\Wset^{1,2}(\Dset,\rho^2)}(\Pbb,\Qbb)$.
+\section{Proofs}
 
-\paragraph{Neighborhood graph.}
+\subsection{Proof of Theorem~\ref{thm:sobolev_testing_rate}}
 
-Let $G_{n,r_n} = (V,E,w)$ denote a weighted, undirected graph constructed from the samples $\Zbf = \set{z_1, \ldots, z_n} = (\Xbf,\Ybf)$ where $V = \set{1, \ldots, n}$, and $w_{uv} = K(z_u, z_v):= k(\frac{\norm{z_u - z_v}}{\epsilon_n}) \geq 0$ for $u,v \in V$, and a particular kernel function $k$.
-Here $(u,v) \in E$ if and only if $w_{uv} > 0$.
-
-Motivated by the integral probability metric, our test statistics will be of the form
-\begin{equation*}
-\gamma_{\F_n}(\Pbb_n,\Qbb_n) = \sup_{f_n \in \F_n}\abs{\int_{D} f_n d\Pbb_n - \int_{D} f_n d\Qbb_n}
-\end{equation*}
-where
-\begin{equation*}
-\Pbb_n := \frac{1}{n_1}\sum_{i = 1}^{n_1} \delta_{x_i}, ~~ \Qbb_n := \frac{1}{n_2}\sum_{i = 1}^{n_2} \delta_{y_i}
-\end{equation*}
-are the empirical distributions of $\Xbf$ and $\Ybf$, respectively, and $\F_n$ is a \textcolor{red}{class of functions $f_n: \Zbf \to \Reals$ exhibiting some regularity with respect to the neighborhood graph $G_{n,r}$.}
-
-\paragraph{Laplacian smoothing and Total Variation denoising.}
-
-For convenience, we number the edges $E  = (e_1, \ldots, e_m)$.
-We denote by $\Bbf \in \Reals^{m \times n}$ the edge incidence matrix of $G_{n,r}$, which for $k$th edge $e_k = (u,v)$ has $k$th row $\Bbf_{k} = (0,\ldots,-w_{uv},\ldots,w_{uv},\ldots,0)$ with a $-w_{uv}$ in the $u$th location, and a $w_{uv}$ in the $v$th location.
-The random (unnormalized) Laplacian matrix is then $\Lbf = \Bbf^T \Bbf$.
-We also introduce a \emph{label vector}, given by $\lbf = (\ell_1, \ldots, \ell_n)$ with
+To prove Theorem~\ref{thm:sobolev_testing_rate}, we will show that when
 \begin{equation}
-\label{eqn: label_vector}
-\ell_k = 
-\begin{cases}
-\frac{n}{n_1},~~~ \text{$z_k \in \Xbf$} \\
--\frac{n}{n_2}, ~ \text{$z_k \in \Ybf$}
-\end{cases}
+\label{eqn:critical_radius}
+\epsilon^2 \geq c \cdot b \cdot (\log n)^{h(\beta,d)} \cdot n^{-4/(4 + d)}
+\end{equation}
+and we choose tuning parameters as specified in Theorem~\ref{thm:sobolev_testing_rate}, the test $\phi_{\mathrm{spec}}$ satisfies
+\begin{equation}
+\mathcal{R}_{\epsilon}(\phi_{\spec}; W^{1,2}(\mathcal{X};R)) = \Ebb_{f_0}(\phi_{\spec}) + \sup_{\substack{f \in W^{1,2}(\mathcal{X};R) \\ \norm{f}_{\Leb^2(\mathcal{X})} > \epsilon}} \Ebb_{f}(1 - \phi_{\spec}) \leq \left(\frac{1}{b^2} + \frac{3}{b}\right) + o(1).
 \end{equation}
 
-Our test statistics $T_1(\lbf; G_{n,r})$ and $T_2(\lbf; G_{n,r})$ are defined as follows:
+Our analysis will proceed according to the following steps.
+\begin{enumerate}
+	\item We upper bound the testing error of our eigenvector projection test statistic when $Y$ are viewed as random responses defined over the vertices of a fixed graph $G$.
+This upper bound will hold whenever certain functionals on the graph $G$ are themselves bounded.
+	\item We analyze the behavior of these functionals with respect to the random graph $G_{n,r}$, and bound them with high probability.
+	\item We condition on $X \subseteq \mathcal{E}$, where $\mathcal{E}$ is a high-probability set over which the relevant functionals on $G_{n,r}$ are bounded.
+We conclude that the upper bound on testing error derived in our first step holds whenever $X \subseteq \mathcal{E}$.
+\end{enumerate}
+
+\subsubsection{Step 1: Testing error on a fixed graph}
+
+Let $G = (V,E)$ be a graph over vertices $V = \set{v_1,\ldots,v_n}$, and let $\beta = (\beta_1,\ldots,\beta_n) \in \Reals^n$ be a signal over the vertices $V$.
+We observe responses $Y = (y_1,\ldots,y_n)$ according to the model
+\begin{equation*}
+y_i = \beta_i + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
+\end{equation*}
+Letting $L = VSV^T$ be the spectral decomposition of the Laplacian $L$ of $G$, our graph spectral test statistic is 
+\begin{equation*}
+T_{\spec} = \frac{1}{n}\sum_{k = 1}^{\kappa} \left(\sum_{i = 1}^{n} v_i y_i\right)^2
+\end{equation*}
+where $\kappa$ is a tuning parameter. The resulting test we will use is
+\begin{equation*}
+\phi_{\spec} = \1\{T_{\spec} \geq \frac{\kappa}{n} + t(b)\},~~\textrm{where}~ t(b) = b\sqrt{\frac{2\kappa}{n^2}}~~\textrm{for $b \geq 1$.}
+\end{equation*}
+
+Let $S_2(\beta;G)$ be a measure of smoothness the signal $\beta$ displays over the graph $G$, given by
+\begin{equation*}
+S_2(\beta;G) = \sum_{i = 1}^{n} \sum_{j = 1}^{n} (\beta_i - \beta_j)^2 \1((v_i,v_j) \in E).
+\end{equation*}
+In Lemma~, we upper bound the Type I and Type II error of the test $\phi_{\spec}$.
+Our bound on the Type II error will be stated as a function of $S_2(\beta;G)$ as well as the $\kappa$th eigenvalue $s_{\kappa}$.
+
+\begin{lemma}
+	\label{lem:fixed_graph_testing}
+	Let $1 \leq \kappa \leq n$ be an integer.
+	\begin{enumerate}
+		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}$ is upper bounded
+		\begin{equation}
+		\label{eqn:graph_spectral_type_I_error}
+		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{1}{b^2}.
+		\end{equation}
+		\item \textbf{Type II error:} For any $b$ and $\beta$ such that
+		\begin{equation}
+		\label{eqn:fixed_graph_testing_critical_radius}
+		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_2(\beta;G)}{ns_{\kappa}}
+		\end{equation}
+		the Type II error of $\phi_{\beta}$ is upper bounded,
+		\begin{equation}
+		\label{eqn:graph_spectral_type_II_error}
+		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{3}{b}.
+		\end{equation}
+	\end{enumerate}
+\end{lemma}
+
+The first term on the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius} grows with $\kappa$, while the second term shrinks.
+These can be thought of as the variance and squared bias terms, respectively, of the error.
+
+\paragraph{Proof of Lemma~\ref{lem:fixed_graph_testing}.}
+
+To prove Lemma~\ref{lem:fixed_graph_testing} we will first compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}$, and then use Chebyshev's inequality to show \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.
+
+\vspace{.2 in}
+
+\textit{Mean of $T_{\mathrm{spec}}$:} Using the notation $\dotp{v}{w} = \sum_{i = 1}^{n} v_iw_i$, we have
 \begin{align*}
-	T_1(\lbf; G_{n,r}) := \sup_{\fbf \in \Reals^n: \norm{\Bbf \fbf}_1 \leq C_{n,r}} ~ \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_k \\
-	T_2(\lbf; G_{n,r}) := \sup_{\fbf \in \Reals^n: \norm{\Bbf \fbf}_2^2 \leq C_{n,r}} ~ \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_k
+\Ebb(T_{\spec}) & = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{\beta}{v_k}^2 + \Ebb\bigl( \dotp{\varepsilon}{v_k}^2 + 2 \dotp{\varepsilon}{v_k} \dotp{\beta}{v_k}\bigr)\right) \\
+& = \frac{\kappa}{n} + \frac{1}{n}\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2.
 \end{align*}
-where $C_{n,r} = \frac{\sigma_k}{n^2 r_n^{d+2}}$ and we write $\fbf = (f_1,\ldots, f_n)$. We note that, as promised, these satisfy the form
+When $\beta = 0$, this equals $\kappa/n$. Otherwise, we have the following lower bound:
 \begin{align*}
-T_1(\lbf; G_{n,r}) & = \sup_{f_n \in TV_n}\abs{\int_{D} f_n d\Pbb_n - \int_{D} f_n d\Qbb_n} \\
-T_2(\lbf; G_{n,r}) & = \sup_{f_n \in \Wset_n}\abs{\int_{D} f_n d\Pbb_n - \int_{D} f_n d\Qbb_n}
+\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2 & = \norm{\beta}_2^2 - \sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \\
+& \geq \norm{\beta}_2^2 - \frac{1}{s_{\kappa}}\sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 s_k \\
+& \geq \norm{\beta}_2^2 - \frac{S_2(\beta;G)}{s_{\kappa}},
 \end{align*}
-where $TV_n = \set{\fbf: \norm{\Bbf \fbf}_1 \leq C_{n,r}}$ and $\Wset_n = \set{\fbf: \norm{\Bbf \fbf}_2 \leq C_{n,r}}$.
+and therefore $\Ebb(T_{\spec}) \geq \kappa/n + n^{-1}(\norm{\beta}_2^2 - S_2(\beta;G)/s_{\kappa})$.
 
-\section{Consistency under fixed alternative}
+\vspace{.2 in}
 
-\paragraph{Binomialized data model.}
-For technical reasons, we would like $z_1, \ldots, z_n$ to be independent and identically distributed. We consider the following generative model, which we term the \emph{binomialized data model}:
+\textit{Variance of $T_{\mathrm{spec}}$:}
+We write $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvectors $v_1,\ldots,v_{\kappa}$ as columns. Consequently,
+\begin{align}
+\Var(T_{\spec}) & = \frac{1}{n^2} \Var(y^T V_{\kappa} V_{\kappa}^T y) \\
+& = \frac{1}{n^2} \Var((\beta + \varepsilon)^T V_{\kappa} V_{\kappa}^T (\beta + \varepsilon)) \\
+& = \frac{1}{n^2} \Var(2 \beta^T V_{\kappa} V_{\kappa}^T \varepsilon + \varepsilon^T V_{\kappa} V_{\kappa}^T \varepsilon) \\
+& \leq \frac{1}{n^2}(4 \beta^T V_{\kappa} V_{\kappa}^T \beta + 2\kappa)
+\end{align}
+where the last inequality follows from standard properties of the Gaussian distribution.
+We now move on to showing the desired inequalities \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.
+For simplicity, we introduce the notation
+\begin{equation*}
+t(b) = b\sqrt{\frac{2\kappa}{n^2}}, ~~ \Delta = \norm{\beta}_2^2 - \frac{S_2(\beta;G)}{s_{\kappa}}.
+\end{equation*}
 
-Fix $n \in \Naturals > 0$, and $n_1 \sim \Bin(n,1/2),~ n_2 = n - n_1$.
-Then, let $x_1, \ldots, x_{n_1} \in \Rd$ be a sequence of i.i.d random points chosen according to $\Pbb$, and $y_1, \ldots, y_{n_2} \in \Rd$ a separate sequence of i.i.d random points chosen according to $\Qbb$, with $x_j \perp y_k$ for all $j,k$.
-Fix $\widetilde{\Zbf} = (\widetilde{z}_1,\ldots,\widetilde{z}_n) := (x_1, \ldots, x_{n_1}, y_1, \ldots, y_{n_2})$.
-Finally, for a permutation $\pi: [n] \to [n]$ chosen uniformly at random among all such permutations, let $\Zbf = (z_1,\ldots,z_n) =  (\widetilde{z}_{\pi(1)}, \ldots, \widetilde{z}_{\pi(n)})$.
+\vspace{.2 in}
 
-The label vector $\lbf$ remains defined as in \eqref{eqn: label_vector} with respect to $\Zbf$.
-Note that now $z_i \overset{i.i.d}{\sim} \frac{\Pbb}{2} + \frac{\Qbb}{2}$, as we desired, with density function $\mu(x) := \frac{p(x) + q(x)}{2}$.
-
-\begin{theorem}[Pointwise limit of Laplacian smooth test statistic.]
-	\label{thm: LLN_for_t2}
-	Let $d \geq 2$ and let $\Dset \subset \Rd$ be an open, bounded, connected set with Lipschitz boundary. Let $\mu$ satisfy
-	\begin{equation*}
-	m \leq \mu(x) \leq M \tag{$\forall x \in D$}
-	\end{equation*}
-	for some $0 < m \leq M$. Let $(r_n)$ be a sequence of positive numbers converging to $0$ and satisfying
+\textit{Proof of~\eqref{eqn:graph_spectral_type_I_error}:} By Chebyshev's inequality,
 \begin{align*}
-	& \lim_{n \to \infty} \frac{(\log n)^{3/4}}{n^{1/2}} \frac{1}{r_n} = 0 ~~ \text{if $d = 2$} \\
-	& \lim_{n \to \infty} \frac{(\log n)^{1/d}}{n^{1/d}} \frac{1}{r_n} = 0 ~~ \text{if $d \geq 3$}
-	\end{align*}
-	Assume the kernel $k$ satisfies conditions: 
-	\begin{align*}
-	& k(0) > 0 ~ \text{and $k$ is continuous at $0$.} \tag{\textbf{K1}} \\
-	& k~ \text{is non-increasing.} \tag{\textbf{K2}} \\
-	& \text{The integral $\int_{0}^{\infty} k(r) r^{d+1} dr$ is finite.} \tag{\textbf{K3}}
+\Pbb_{\beta = 0}\bigl(T_{\spec} \geq \frac{\kappa}{n} + t(b)\bigr)
+& \leq \Pbb_{\beta = 0}\bigl(\abs{T_{\spec} - \frac{\kappa}{n}} \geq t(b)\bigr) \\
+& \leq \frac{\Var_{\beta = 0}(T_{\spec})}{t(b)^2} = \frac{1}{b^2}.
 \end{align*}
 
-	Then with probability one the following statement holds: For $(z_1, \ldots, z_n)$ chosen under the binomialized data model,
+\vspace{.2 in}
+
+\textit{Proof of~\eqref{eqn:graph_spectral_type_II_error}:} Assumption~\eqref{eqn:fixed_graph_testing_critical_radius} implies
 \begin{equation*}
-	\lim_{n \to \infty} T_2(\lbf; G_{n,r_n}) = \gamma_{\Wset^{1,2}(\Dset,\mu^2)}(\Pbb,\Qbb).
+\Delta \geq \frac{\norm{\beta}_2^2}{n} - \frac{C_n^2}{nC^2} \geq 2 t(b).
+\end{equation*}
+Then another application of Chebyshev's inequality gives us
+\begin{align*}
+\Pbb_{\beta}\bigl(T_{\spec} \leq \frac{\kappa}{n} + t(b)\bigr) & = \Pbb_{\beta}\bigl(T_{\spec} - \Ebb_{\beta}(T_{\spec}) \leq t(b) - \Delta \bigr) \\
+& \leq \Pbb_{\beta}\bigl(\abs{T_{\spec} - \Ebb_{\beta}(T_{\spec})} \leq \Delta - t(b) \bigr) \tag{since $\Delta \geq t(b)$}	\\
+& \leq \frac{\Var_{\beta}(T_{\spec})}{(\Delta - t(b))^2} \\
+& \leq 4\frac{\Var_{\beta}(T_{\spec})}{\Delta^2} \tag{since $\Delta \geq 2t(b)$} \\
+& \leq 4\frac{2\kappa/n^2 + 4\beta^T V_{\kappa} V_{\kappa}^T \beta /n}{\Delta^2}.
+\end{align*}
+We handle each summand in the numerator separately. For the first term, since $\Delta \geq 2 t(b)$, we have
+\begin{equation}
+\label{eqn:spectral_type_II_error_pf1}
+\frac{2\kappa}{n^2\Delta^2} \leq \frac{1}{2b^2}.
+\end{equation}
+
+For the second term, noting that $\Delta = \beta^T V_{\kappa} V_{\kappa}^T \beta$, we have
+\begin{align}
+\frac{\beta^T V_{\kappa} V_{\kappa}^T \beta/n}{\Delta^2} & \leq \frac{1}{n\Delta} \nonumber \\
+& \leq \frac{1}{2nt(b)} \nonumber \\
+& = \frac{1}{2b\sqrt{2\kappa}}, \label{eqn:spectral_type_II_error_pf2}
+\end{align}
+and combining~\eqref{eqn:spectral_type_II_error_pf1} and~\eqref{eqn:spectral_type_II_error_pf2} yields~\eqref{eqn:graph_spectral_type_II_error}.
+
+\subsubsection{Step 2: Bounding neighborhood graph functionals}
+
+To make use of Lemma~\ref{lem:fixed_graph_testing} we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$.
+In particular, we will show that for some constants $c_1,c_2,c_3$, the following statements:
+\begin{enumerate}[label=(E\arabic*)]
+	\item 
+	\label{event:discrete_sobolev_norm}
+	\textbf{Graph Sobolev norm:} When $\beta = (f(x_1),\ldots,f(x_n))$ for $f \in W^{1,2}(\mathcal{X};R)$:
+	\begin{equation}
+	\label{eqn:continuous_to_discrete_sobolev_norm}
+	S_2(\beta;G_{n,r}) \leq c_1 \cdot n^2 r^{d + 2} 
+	\end{equation}
+	\item 
+	\label{event:eigenvalue_tail_decay}
+	\textbf{Eigenvalue tail bound:} If $r = (\log^{a}n)\cdot(\log n/n)^{1/d}$ for any $a > 0$, then for all $\kappa \in [n]$:
+	\begin{equation}
+	\label{eqn:eigenvalue_tail_bound}
+	s_{\kappa} \geq c_2 \cdot \frac{\kappa^{2/d}}{n^{2/d}}
+	\end{equation}
+	\item 
+	\label{event:l2_norm}
+	\textbf{Empirical norm of $f$:} When $f \in W^{1,2}(\mathcal{X};R)$ and one of (a) $d = 1,2$, or (b) $d = 3$ and $\norm{f}_{\Leb^2}^2 \geq n^{-2/(d - 2)}$, then
+	\begin{equation}
+	\label{eqn:l2_to_empirical_norm}
+	\frac{1}{n}\sum_{i = 1}^{n}\beta_i^2 = \norm{f}_n^2 \geq c_3 \cdot \norm{f}_{\Leb^2}^2
+	\end{equation}
+\end{enumerate}
+
+each hold with probability $1 - o(1)$ as $n \to \infty$.
+
+\paragraph{Proof of~\eqref{eqn:continuous_to_discrete_sobolev_norm}:}
+
+Observe that
+\begin{align*}
+\frac{1}{n^2}\Ebb(\beta^T L \beta) = \Ebb\left(\frac{1}{n^2} \sum_{i,j = 1}^{n} (\beta_i - \beta_j)^2 A_{ij}\right) & = \frac{(n - 1)}{n} \int_{\Xset} \int_{\Xset} (f(x) - f(y))^2K_r(x,y) \,dP(y) \,dP(x) \\
+& \leq p_{\max}^2 \frac{(n - 1)}{n} \int_{\Xset} \int_{\X} (f(x) - f(y))^2K_r(x,y) \,dy \,dx.
+\end{align*}
+We will show that for any $f \in \mathcal{W}^{1,2}(\X;R)$
+\begin{equation}
+\label{eqn:nonlocal_continuous_sobolev_norm}
+\int_{\Xset} \int_{\Xset} (f(x) - f(y))^2K_r(x,y) \,dy \,dx \leq c r^{d + 2} \norm{f}_{W^{1,2}(\mathcal{X})}
+\end{equation}
+whence the desired result~\eqref{eqn:continuous_to_discrete_sobolev_norm} follows by Markov's inequality and the assumption $f \in W^{1,2}(\mathcal{X};R)$.
+
+We begin by dealing with complications due to smoothing at the boundary of $\mathcal{X} = [0,1]^d$.
+Let $U$ be any bounded open set such that $\Xset \subset U$ is compactly contained in $U$.
+Note that as $\partial \Xset$ is $C^1$, by Theorem \ref{thm:evans_extension} there exists $g \in W^{1,2}(\Reals^d)$ such that
+\begin{enumerate}[(i)]
+	\item
+	\label{eqn:evans_extension_1}
+	$g = f$, $P$-almost-everywhere in $\Xset$
+	\item 
+	$g$ has support within $V$, and  
+	\item 
+	\label{eqn:sobolev_1_bound_2}
+	$\norm{g}_{W^{1,2}(\Rd)} \leq c \norm{f}_{W^{1,2}(\Xset)}$ for a constant $c$ which depends only on $\Xset$.
+\end{enumerate}
+As a result of \ref{eqn:evans_extension_1},
+\begin{equation}
+\label{eqn:sobolev_1_bound_1}
+\int_{\Xset} \int_{\Xset} (f(x) - f(y))^2K_r(x,y) \,dy \,dx \leq \int_{\Rd} \int_{\Rd} (g(x) - g(y))^2K_r(x,y) \,dy \,dx.
+\end{equation}
+
+Next we smooth $g$, so that we may work with ordinary partial derivatives.
+We let $\eta \in C^{\infty}(\Rd)$ be given by
+\begin{equation*}
+\eta(x) :=
+\begin{cases}
+C \exp \left\{\frac{1}{\norm{x}^2 - 1}\right\}& \quad \textrm{if $\norm{x}_2 \leq 1$} \\
+0 & \quad \textrm{if $\norm{x}_2 \geq 1$}
+\end{cases}
+\end{equation*}
+where the normalizing constant $C > 0$ is chosen so that $\int_{\Rd} \eta dx = 1$. Let $\eta_r(x) := (1/r^d) \eta(x/r)$. Then, the mollification of $g$ by $\eta_r$ is given by
+\begin{align*}
+g^r & := g \ast \eta_r \\
+& = \int_{\Rd} \eta_r(x - y) g(y) dy
+\end{align*}
+(Refer to \textcolor{red}{Evans}, Appendix C, Theorem 7 for a proof that $g^r \in C^{\infty}(\Rd)$.)
+Adding and subtracting within \eqref{eqn:sobolev_1_bound_1}, we have
+\begin{align}
+\int_{\Rd} & \int_{\Rd} (g(x) - g(y))^2K_r(x,y) \,dy \,dx \\
+& \leq 3 \int_{\Rd} \int_{\Rd} \bigl((g(x) - g^r(x))^2 + (g^r(x) - g^r(y))^2 + (g^r(y) - g(y))^2\bigr)K_r(x,y) \,dy \,dx \nonumber \\
+& = 6 \int_{\Rd} \int_{\Rd} \bigl((g(y) - g^r(y))^2 K_r(x,y) \,dy \,dx + 2 \int_{\Rd} \int_{\Rd} \bigl((g^r(x) - g^r(y))^2 K_r(x,y) \,dy \,dx \label{eqn:sobolev_1_bound_5}
+\end{align}
+We deal with each summand individually, beginning with the first one. Let $\nu_d = \pi^{d/2}/\Gamma(d/2 + 1)$ denote the volume of a ball with unit radius in $\Rd$. Then, we have
+\begin{align}
+\int_{\Rd} \int_{\Rd} \bigl((g(y) - g^r(y))^2 K_r(x,y) \,dy \,dx & = \int_{\Rd} \int_{B(x,r)} \bigl((g(y) - g^r(y))^2\,dy \,dx \nonumber \\
+& = \int_{\Rd} \int_{B(y,r)} \bigl((g(y) - g^r(y))^2\,dx \,dy \tag{by Tonelli's Theorem} \nonumber \\
+& = \nu_d r^d \int_{\Rd}\bigl((g(y) - g^r(y))^2 \,dy \\
+& \leq \nu_d r^{d + 2} \int_{\Rd} \norm{\nabla g(y)}^2 \,dy \label{eqn:sobolev_1_bound_4}
+\end{align}
+where the last line follows from Lemma \ref{lem:poincare_mollify}.
+
+We now turn our attention to the second summand. Note that as $g^r \in C^{\infty}(\Rd)$, we may apply Theorem \ref{thm:taylor_expansion} and obtain
+\begin{align*}
+(g^r(x) - g^r(y))^2 & = \left(\int_{0}^{1} \nabla g^r(x + t(y - x)) \cdot (y - x) \,dt\right)^2 \\
+& \leq \int_{0}^{1} \bigl(\nabla g^r(x + t(y - x)) \cdot (y - x)\bigr)^2 \,dt  \tag{Jensen's inequality} \\
+& \leq \norm{y - x}^2 \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 \,dt. \tag{Cauchy-Schwarz inequality}
+\end{align*}
+As a result, we have
+\begin{align*}
+\int_{\Rd} \int_{\Rd} \bigl((g^r(x) - g^r(y))^2 K_r(x,y) \,dy) \,dx & \leq \int_{\Rd} \int_{\Rd} \norm{y - x}^2  \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 K_r(x,y) \,dt \,dy \,dx \\
+& \leq r^{2} \int_{\Rd} \int_{\Rd} \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 K_r(x,y) \,dt \,dy \,dx \\
+& = r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + t(y - x))}^2 K_r(x,y) \,dx \,dt \,dy \\
+& = r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + tz)}^2 K_r(z) \,dx \,dt \,dz \tag{$z = y - x$}
+\end{align*}
+
+where we write $K_r(z) = \mathbf{1}(\norm{z} \leq r)$ in an abuse of notation. Next, we note that
+\begin{equation*}
+\int_{\Rd} \norm{\nabla g^r(x + tz)}^2 \,dx = \int_{\Rd} \norm{\nabla g^r(x)}^2 \,dx \leq \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
+\end{equation*}
+with the inequality following from Lemma \ref{lem:gradient_mollify_commute}. Therefore,
+\begin{align}
+r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + tz)}^2 K_r(z) \,dx \,dt \,dz & \leq r^{2} \int_{\Rd} K_r(z) \int_{0}^{1} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx  \,dt \,dz \nonumber \\
+& = \nu_d r^{2 + d} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx. \label{eqn:sobolev_1_bound_3}
+\end{align}
+By \eqref{eqn:sobolev_1_bound_5}, \eqref{eqn:sobolev_1_bound_4} and \eqref{eqn:sobolev_1_bound_3}, we have that 
+\begin{equation*}
+\int_{\Rd} \int_{\Rd} (g(x) - g(y))^2 K_r(x,y) \,dx \,dy \leq 9 \nu_d r^{d + 2} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
+\end{equation*}
+Then by \ref{eqn:sobolev_1_bound_2}, $\int_{\Rd} \norm{\nabla g(x)}^2 \,dx \leq c \int_{\Xset} \norm{\nabla f(x)}^2 \,dx$, and the desired result of \eqref{eqn:sobolev_1_bound_1} follows.
+
+\paragraph{Proof of~\eqref{eqn:eigenvalue_tail_bound}:}
+
+For ease of notation we write $t = n^{1/d}$, and assume without loss of generality that $t$ is whole number. Let $\overline{G} = (\overline{V},\overline{E})$ be the lattice over $[0,1]^d$, formally
+\begin{align*}
+\overline{V} & = \bigl\{\overline{x}_k := \frac{1}{n}(k_1,\ldots,k_d): k \in [t]^d\bigr\} \\
+\overline{E} & = \bigl\{(\overline{x}_k, \overline{x}_{\ell}): \norm{k - \ell}_1 = 1\bigr\}.
+\end{align*}
+
+By Theorem 1.1 of \textcolor{red}{Garcia Trillos and Slepcev}, with probability at least $1 - n^{-1}$ there exists a bijection $\pi: \overline{V} \to X$ such that
+\begin{equation*}
+\max_{k \in [t]^d} \abs{\overline{x}_k - \pi(\overline{x}_k)} \leq c \left(\frac{\log n}{n}\right)^{1/d}
+\end{equation*}
+Since our choice $r = \log^a(n) \cdot (\log n/n)^{1/d}$ implies $r = \omega(\left(\frac{\log n}{n}\right)^{1/d})$, for $n$ sufficiently large with probability $1 - n^{-1} = 1 - o(1)$ we have
+\begin{equation}
+\label{eqn:eigenvalue_tail_bound_pf1}
+(\overline{x}_k,\overline{x}_{\ell}) \in \overline{E} \Longrightarrow (\pi(\overline{x}_k),\pi(\overline{x}_{\ell})) \in E.
+\end{equation}
+Let $\overline{L}$ be the Laplacian of $\overline{G}$, with eigenvalues $\lambda_1 < \cdots \leq \lambda_n$. If \eqref{eqn:eigenvalue_tail_bound_pf1} holds, then for all $x \in \Reals^n$,
+\begin{equation*}
+x^T \overline{L} x \geq x^T L x \Longrightarrow \lambda_j \leq s_j~~ \textrm{for all}~ j \in [n].
+\end{equation*}
+The eigenvalues of the $d$-dimensional lattice are explicitly known.
+In particular when $j = h^d$ for some whole number $h$, we have $s_j \geq \lambda_j = 4 d \sin^2(\pi h/2t) \geq \pi^2 h^2/ (4 t^2)$, with a similar inequality (for possible different constants) holding when $j^{1/d}$ is not integral; this shows~\eqref{eqn:eigenvalue_tail_bound}.
+
+\paragraph{Proof of~\eqref{eqn:l2_to_empirical_norm}:}
+
+To prove~\eqref{eqn:l2_to_empirical_norm} we will upper bound
+\begin{equation*}
+\mathbb{E}\biggl[\Bigl(\frac{1}{n}\sum_{i = 1}^{n}f^2(x_i)\Bigr)^2\biggr],
+\end{equation*}
+and then apply the Paley-Zygmund inequality (Lemma~\ref{lem:paley_zygmund}). We have
+\begin{align*}
+\mathbb{E}\biggl[\Bigl(\frac{1}{n}\sum_{i = 1}^{n}f^2(x_i)\Bigr)^2\biggr] & = \frac{(n-1)}{n}\mathbb{E}(f^2(x_1)) + \frac{\mathbb{E}[f^4(x_1)]}{n} \\
+& \leq p_{\max}^2 \left(\norm{f}_{\Leb^2}^2 + \frac{\norm{f}_{\Leb^4}^4}{n}\right) \\
+& \overset{(i)}{\leq} p_{\max}^2 \left(\norm{f}_{\Leb^2}^2 + c\frac{\norm{f}_{\Leb^2}^{(4 - d)}}{n}\right) \\
+& \overset{(ii)}{\leq}  p_{\max}^2 \left(\norm{f}_{\Leb^2}^2 + c\norm{f}_{\Leb^2}^2\right).
+\end{align*}
+Here $(i)$ follows from two facts, the first being the embedding $W^{1,2}(\mathcal{X};R) \subseteq \Leb^q(\mathcal{X};R')$ for $q = 2d/(d - 2)$ when $d > 2$ and $q = \infty$ when $d \leq 2$ and some constant $R'$, and the second being a standard result on interpolation of $\Leb^p$ spaces, namely that since $f \in \Leb^2(\mathcal{X}) \cap \Leb^q(\mathcal{X})$, then
+\begin{equation*}
+\norm{f}_{\Leb^4} \leq \norm{f}_{\Leb^2}^{\lambda} \cdot \norm{f}_{\Leb^q}^{1 - \lambda},~~\textrm{where}~ \frac{1}{4} = \frac{\lambda}{2} + \frac{1 - \lambda}{q}.
+\end{equation*} 
+Then $(ii)$ follows from our assumptions on $\norm{f}_{\Leb^2}^2$: either 
+\begin{enumerate}[(a)]
+	\item $d \leq 2$ and $\norm{f}_{\Leb^2}^2 \geq 1/n$ so $\norm{f}_{\Leb^2}^{(4 - d)} \leq \max(R^{4 - d}/n, \norm{f}_{\Leb^2}^2) \leq c\norm{f}_{\Leb^2}^2$, or
+	\item $d = 3$ and $\norm{f}_{\Leb^2}^2 \geq n^{-2/(d - 2)}$.
+\end{enumerate} 
+Applying Lemma~\ref{lem:paley_zygmund} to  $\norm{f}_n^2$ yields~\eqref{eqn:l2_to_empirical_norm}.
+
+\subsubsection{Step 3: Conclusion}
+
+We note that for all possible values of $X \in \Xset^n$, under the null hypothesis $f = f_0 = 0$ and therefore $\beta = (f(x_1),\ldots,f(x_n)) = 0$ as well.
+Therefore by~\eqref{eqn:graph_spectral_type_I_error}, we have the following bound on Type I error:
+\begin{equation}
+\Ebb_{f_0}(\phi_{\mathrm{spec}}) = \mathbb{E}(\mathbb{E}_{\beta = 0}(\phi_{\spec}) | X) \leq \frac{1}{b^2}.
+\end{equation}
+
+Now, we bound Type II error under the assumption $f \in W^{1,2}(\mathcal{X};R)$ and 
+\begin{equation}
+\label{eqn:critical_radius_1}
+\norm{f}_{\Leb^2}^2 \geq \epsilon^2 = c \cdot b \cdot (\log n)^{h(\beta,d)} \cdot n^{-4/(4 + d)}.
+\end{equation}
+We verify that the conditions required to apply Step 2 are satisfied.
+In particular, we choose $r = (\log^a n) \cdot (\log n/n)^{1/d}$, and observe that~\eqref{eqn:critical_radius_1} implies $\norm{f}_{\Leb^2}^2 \geq n^{-1}$ when $d \leq 2$ and $\norm{f}_{\Leb^2}^2 \geq n^{-2/(d - 2)}$ when $2 < d < 4$.
+We may therefore apply our conclusions in Step 2; namely, that for every possible choice of $f$ there exists a good set $\mathcal{E}_f \subseteq \Xset^n$ with $\Pbb(\mathcal{E}_f) \geq 1 - o(1)$ such that each of \eqref{eqn:continuous_to_discrete_sobolev_norm}, \eqref{eqn:eigenvalue_tail_bound}, and \eqref{eqn:l2_to_empirical_norm} hold for all $X \subseteq \mathcal{E}_f$.
+Choosing $\kappa = n^{2d/(4 + d)}$ to balance the squared bias and variance terms on the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius}, we have that for all $X \subseteq \mathcal{E}_f$
+\begin{align*}
+2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_2(\beta;G)}{ns_{\kappa}} & \leq 2bn^{-4/(4+d)} + \frac{n r^{d + 2}}{s_{\kappa}} \tag{by \eqref{eqn:continuous_to_discrete_sobolev_norm}} \\
+& \leq 2bn^{-4/(4+d)} + c\cdot\frac{n r^{d + 2}}{n^{4/(4+d)}} \tag{by \eqref{eqn:eigenvalue_tail_bound}} \\
+& \leq 2bn^{-4/(4+d)} + c\cdot(\log^{a(d + 2) + (1 + 2/d)}n)\cdot n^{-4/(4+d)} \\
+& \leq \frac{1}{10}\norm{f}_{\Leb^2} \tag{for a suitably large choice of $c$ in \eqref{eqn:critical_radius_1}}\\
+& \leq \frac{1}{n}\sum_{i = 1} \beta_i^2. \tag{by \eqref{eqn:l2_to_empirical_norm}}
+\end{align*}
+We conclude that for all $X \subseteq \mathcal{E}_f$, the inequality \eqref{eqn:fixed_graph_testing_critical_radius} is satisfied with respect to $\beta = (f(x_1),\ldots,f(x_n))$ and $G = G_{n,r}$.
+As a result the worst-case Type II error is bounded
+\begin{equation*}
+\sup_{\substack{f \in W^{1,2}(\mathcal{X;R}) \\ \norm{f}_{\Leb^2} \geq \epsilon}}\mathbb{E}_{f}(1 - \phi_{\spec}) \leq \sup_{\substack{f \in W^{1,2}(\mathcal{X;R}) \\ \norm{f}_{\Leb^2} \geq \epsilon}} \mathbb{E}\bigl[\mathbb{E}_{\beta}(1 - \phi_{\spec}|X \in \mathcal{E}_f)\bigr] + o(1) \leq \frac{3}{b} + o(1).
+\end{equation*}
+
+\subsection{Proof of Proposition~\ref{prop:L4_testing_rate}}
+
+Let $T_{\mathrm{mean}} = \frac{1}{n}\sum_{i = 1}^{n} y_i^2$. The expectation of $T_{\mathrm{mean}}$ is
+\begin{equation*}
+\Ebb(T_{\mathrm{mean}}) = \mathbb{E}(f^2(x_1)) + 1,
+\end{equation*}
+and the variance can be upper bounded
+\begin{equation*}
+\Var(T_{\mathrm{mean}}) \leq \frac{1}{n}(3 + p_{\max} \norm{f}_{\Leb^4}^4 + p_{\max}\norm{f}_{\Leb^2}^2) = \frac{c}{n}.
+\end{equation*}
+Since $\mathbb{E}(f^2(x_1)) \geq p_{\min} \epsilon^2$, when $\epsilon^2 \gtrsim b\cdot n^{-1/2}$ we can apply Chebyshev's inequality to obtain the claimed result.
+\section{Auxiliary Results}
+
+\begin{theorem}[\textcolor{red}{Evans} Chapter 5.4, Theorem 1]
+	\label{thm:evans_extension}
+	Assume $U$ is bounded and $\partial U$ is $C^1$.
+Select a bounded open set $V$ such that $U \subset \subset V$ ($U$ is compactly contained in $V$).
+Then there exists a bounded linear operator $E: W^{1,2}(U) \to W^{1,2}(\Rd)$ such that for each $u \in W^{1,2}(U)$:
+	\begin{enumerate}
+		\item $Eu = u$ a.e. in $U$,
+		\item $Eu$ has support within $V$, and 
+		\item 
+		\begin{equation*}
+		\norm{Eu}_{W^{1,2}(\Rd)} \leq C \norm{u}_{W^{1,2}(\Rd)}
+		\end{equation*}
+		the constant $C$ depending only on $U$ and $V$.
+	\end{enumerate}
+\end{theorem}
+
+For $u \in W^{1,2}(\Rd)$ and $x \in \Rd$, write $\nabla u(x) = (D^{e_1}(x),\ldots,D^{e_d}(x)$, for the gradient of $u$.
+
+\begin{theorem}[\textcolor{red}{Evans} Chapter 5.8.1, Theorem 2]
+	\label{thm:evans_poincare}
+	There exists a constant $C$, depending only on $d$, such that
+	\begin{equation*}
+	\norm{u - u^r}_{L^2(B(x,r))} \leq Cr\norm{\nabla u}_{L^2(B(x,r))}
 	\end{equation*}
@@
 
-\section{Proofs}
-
-Fix $\mu_n = \frac{d\Pbb_n + d\Qbb_n}{2}$. \textcolor{red}{We will show a variational form of convergence of $\mu_n$ to $\mu$.}
-
-\subsection{Gamma convergence of constraint}
-
-\begin{definition}[$TL^2$ convergence]
-	Denote by $\mathfrak{B}(\Dset)$ the Borel $\sigma$-algebra of $\Dset$ and $\mathcal{P}(\Dset)$ the set of all Borel probability measures on $\Dset$.
-Given a Borel map $T: \Dset \to \Dset$, the \emph{push-forward} of $\mu$ by $T$ is given by
+\begin{theorem}[Taylor expansion.]
+	\label{thm:taylor_expansion}
+	For any function $u \in C^1$, and any $x,y \in \Rd$, 
 	\begin{equation*}
-	T_{\star}\mu(\Aset) := \mu(T^{-1}(\Aset)), ~~ \Aset \in \mathfrak{B}(\Dset).
+	u(y) - u(x) = \int_{0}^{1} \nabla(u(x + t(y - x))) \cdot (y - x) \,dt
 	\end{equation*}
-
-	Given $\widetilde{\mu} \in \mathcal{P}(\Dset)$, we say that $T$ is a \emph{transportation map} between $\mu$ and $\widetilde{\mu}$ if $T_{\star}\mu = \widetilde{\mu}$.
-If for a sequence $(T_n)$ of transportation maps
-	\begin{equation*}
-	\int_{\Dset} \abs{x - T_n(x)}^2 d\mu(x) \to 0,~ \text{as $n \to \infty$}
-	\end{equation*}
-	we refer to the sequence as \emph{stagnating.}
-
-	Take $f \in L^2(\mu)$ and a sequence $(f_n)$ with $f_n \in L^2(\mu_n)$ for $n = 1,2,\ldots$. If there exists a stagnating sequence of transportation maps $(T_n)$ such that
-	\begin{equation*}
-	\int_{\Dset} \abs{ f - T_n(f_n(x)) }^2 d\mu(x) \to 0,~ \text{as $n \to \infty$}
-	\end{equation*}
-	we say that $(f_n)$ converges $TL^2$ to $f$, and write $f_n \overset{TL^2}{\to} f$.
-\end{definition}
-
-We restate Theorem 1.4 of \cite{trillos2018}, changing notation to match the rest of this paper.
-
-\begin{theorem}[Theorem 1.4 of \cite{trillos2018}]
-	\label{thm: gamma_convergence_energy_functional}
-	Under the setup and conditions of Theorem \ref{thm: LLN_for_t2}, with probability one the following statements hold:
-	\begin{itemize}
-		\item \emph{Liminf inequality:}
-		For all $f \in L^2(\mu)$ and all sequences $(f_n)$ with $f_n \in L^2(\mu_n)$ and $f_n \overset{TL^2}{\to} f$,
-		\begin{equation*}
-		\liminf_{n \to \infty} \frac{1}{n^2 r_n^{d + 2}} \norm{Bf}_2^2 \geq \textcolor{red}{\sigma_{k}} \norm{f}_{1,2,\rho^2}
-		\end{equation*}
-
-		\item \emph{Limsup inequality:} For all $f \in L^2(\mu)$, there exists a sequence $(f_n)$ with $f_n \in L^2(\mu_n)$ and $f_n \overset{TL^2}{\to} f$ such that
-		\begin{equation*}
-		\limsup_{n \to \infty} \frac{1}{n^2 r_n^{d + 2}} \norm{\Bbf \fbf_n }_2^2 \leq \textcolor{red}{\sigma_{k}} \norm{f}_{1,2,\rho^2}
-		\end{equation*}
-		where $\fbf_n = (f_n(z_1), \ldots, f_n(z_n))$.
-
-		\item \emph{Compactness property:} Every sequence $(f_n)$ with $f_n \in L^2(\mu_n)$ satisfying
-		\begin{equation*}
-		\sup_{n \in \Naturals} \frac{1}{n^2 r_n^{d + 2}} \norm{Bf}_2^2 < \infty
-		\end{equation*}
-		is precompact in $TL^2$, that is, every subsequence of $(f_n)$ has a further subsequence which converges in the $TL^2$-sense to an element of $L^2(\Dset)$.
-	\end{itemize}
-
 \end{theorem}
 
-\subsection{Continuity of risk functional}
+\begin{lemma}
+	\label{lem:poincare_mollify}
+	For any function $g \in W^{1,2}(\Rd)$ compactly supported in a bounded open set $V \subset \Rd$, we have
+	\begin{equation}
+	\label{eqn:poincare_mollify}
+	\int_{\Rd} (g(x) - g^r(x))^2 \,dx \leq r^2 \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
+	\end{equation}
+\end{lemma}
+\begin{proof}
+	This Lemma is essentially a reproduction of part of the proof of the Rellich-Kondrachov Compactness Theorem from \textcolor{red}{Evans}.
+Note that it is sufficient to prove in the case when $g$ is smooth.
+To see this, for the moment assume \eqref{eqn:poincare_mollify} holds for all $u \in C^{\infty}(V)$, and let $g \in W^{1,2}(V)$.
+By Theorem \ref{thm:local_approx_smooth_functions}, we may take a sequence $g_m \in C^{\infty}(V)$ such that
+	\begin{equation*}
+	\norm{g - g_m}_{L^2(V)} \to 0, \quad \textrm{and} \quad \norm{\nabla g - \nabla g_m}_{L^2(V)} \to 0,
+	\end{equation*}
+	Then we have
+	\begin{align*}
+	\int_{\Rd} (g(x) - g^r(x))^2 \,dx & \leq \int_{\Rd} (g(x) - g_m(x))^2 \,dx + \int_{\Rd}(g_m(x) - g_m^r(x))^2 \,dx + \int_{\Rd}(g_m^r(x) - g^r(x))^2 \,dx \\
+	& \leq \int_{\Rd} (g(x) - g_m(x))^2 \,dx + r^2 \int_{\Rd} \norm{\nabla g_m(x)}^2 \,dx + \int_{\Rd}(g_m^r(x) - g^r(x))^2 \,dx
+	\end{align*}
+	and taking the limit as $m$ goes to infinity, the right hand side converges to $\int_{\Rd} \norm{\nabla g(x)}^2 \,dx$.
+
+	It remains to show \eqref{eqn:poincare_mollify} in the case where $g$ is smooth. In this case, we have
+	\begin{align*}
+	g^{r}(x) -  g(x) & = \frac{1}{r^d} \int_{B(x,r)} \kappa \left(\frac{x - z}{r}\right)\bigl(g(z) - g(x)\bigr) \,dz \\
+	& = \int_{B(0,1)} \kappa(y) \bigl(g(x - ry) - g(x)\bigr) \,dy \\
+	& = \int_{B(0,1)} \kappa(y) \int_{0}^{1} \frac{d}{dt} \bigl(g(x - try) \bigr) \,dt \,dy \\
+	& = -r \int_{B(0,1)} \kappa(y) \int_{0}^{1} \bigl(\nabla g (x - try) \bigr) \cdot y \,dt \,dy.
+	\end{align*} 
+	Therefore, by Jensen's and Cauchy-Schwarz inequalities, we have
+	\begin{align*}
+	\int_{\Rd} (g(x) - g^r(x))^2 \,dx & \leq r^2 \int_{\Rd} \int_{B(0,1)} \kappa(y) \int_{0}^{1} \norm{\nabla g (x - try)}^2 \norm{y}^2 \,dt \,dy \,dx \\
+	& \leq r^2 \int_{B(0,1)} \kappa(y) \int_{0}^{1} \int_{\Rd} \norm{\nabla g (x - try)}^2 \,dx \,dt \,dy \\
+	& = r^2 \int_{\Rd} \norm{\nabla g(z)}^2 \,dz
+	\end{align*}
+\end{proof}
+
+The following theorem is Theorem 1 in Section 5.3 of \textcolor{red}{Evans}.
+\begin{theorem}[Local approximation by smooth functions.]
+	\label{thm:local_approx_smooth_functions}
+	Assume $U$ is bounded, and $u \in W^{k,p}(U)$ for some $1 \leq p < \infty$. Then there exists functions $u_m \in C^{\infty}(U) \cap W^{k,p}(U)$ such that
+	\begin{equation*}
+	\norm{u_m - u}_{W^{k,p}(U)} \overset{m}{\to} 0.
+	\end{equation*}
+\end{theorem}
 
 \begin{lemma}
-	\label{lem: continuity_of_risk_functional}
-	With probability one the following statement holds: If a sequence $(f_n)$ where $f_n: \set{z_1, \ldots,z_n} \to [-1,1]^{n}$ converges $TL^2$ to $f \in L^2(\mu)$, then
+	\label{lem:gradient_mollify_commute}
+	For any $u \in W^{1,2}(\Rd)$, we have
 	\begin{equation*}
-	\lim_{n \to \infty} \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_n(z_k) = \int_{\Dset} f d\Pbb(x) - \int_{\Dset} fd\Qbb(x).
+	\int_{\Reals^d} \norm{\nabla u^r(x)}_2^2 \,dx \leq \int_{\Reals^d} \norm{\nabla u(x)}_2^2 \,dx
 	\end{equation*}
 \end{lemma}
+\begin{proof}
+	Observe that $\norm{\nabla u(x)}_2^2 = \sum_{j = 1}^{d} (D^{e_j}u(x))^2 $. Therefore,
+	\begin{align*}
+	\int_{\Rd} \norm{\nabla u^r(x)}_2^2 \,dP(x) & = \sum_{j = 1}^{d} \int_{\Rd} (D^{e_j}u^r(x))^2 \,dx \\
+	& = \sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u)^r(x))^2 \,dx
+	\end{align*}
+	where the second equality follows from equation (1) in Section 5.3 of \textcolor{red}{Evans}. Then, for any $v \in L^2(\Rd)$, we have that
+	\begin{align*}
+	\abs{v^r(x)} & = \int_{\Rd} \kappa_r^{1/2}(x - y) \kappa_r^{1/2}(x - y) v(y) \,dy \\
+	& \leq \left(\int_{\Rd} \kappa_r^(x - y) \,dy\right)^{1/2} \left(\int_{\Rd} \kappa_r(x - y) v^2(y) \right)^{1/2} \\
+	& = \left(\int_{\Rd} \kappa_r(x - y) v^2(y) \right)^{1/2}
+	\end{align*}
+	and therefore
+	\begin{align*}
+	\int_{\Rd} (v^r(x))^2 \,dx & \leq \int_{\Rd} \int_{\Rd } \kappa_r(x - y) v^2(y) \,dy \,dx \\
+	& = \int_{\Rd} \int_{\Rd } \kappa_r(x - y) v^2(y) \,dy \,dx \\
+	& = \int_{\Rd} v^2(y) \,dy
+	\end{align*}
 
-\subsection{Proof of Theorem \ref{thm: LLN_for_t2}}
+	Applying this to $D^{e_j}u \in L^2(\Rd)$, we have that
+	\begin{equation*}
+	\sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u)^r(x))^2 \,dx \leq \sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u(x))^2 \,dx = \int_{\Rd} \norm{\nabla u(x)}^2 \,dx.
+	\end{equation*}
+\end{proof}
 
-Let $f^{\star}$ be the witness function for $\gamma_{\Wset^{1,2}(\Dset,\mu^2)}(\Pbb,\Qbb)$, meaning
-\begin{equation*}
-\abs{\int_{\Dset}f^{\star}d\Pbb - \int_{\Dset}f^{\star}d\Qbb} = \gamma_{\Wset^{1,2}(\Dset,\mu^2)}(\Pbb,\Qbb)
-\end{equation*}
-where $f^{\star} \in \Wset^{1,2}(\Dset,\mu^2)$ implies $f^{\star} \in L^2(\mu)$.
-By the limsup inequality in Theorem \ref{thm: gamma_convergence_energy_functional}, there exists some $(f_n) \overset{TL^2}{\to} f$ with $f_n \in L^2(\mu_n)$ such that
+\begin{lemma}
+	\label{lem:paley_zygmund}
+	Let $0 < p < q$, and let $Z$ be a non-negative random variable such that $\mathbb{E}(Z^q) < \infty$. Then, for all $0 \leq \lambda \leq 1$,
 	\begin{equation}
-\label{eqn: limsup_discrete_energy}
-\limsup_{n \to \infty} \frac{1}{n^2 r_n^{d+2}} \norm{\Bbf \fbf_n}_2^2 \leq \textcolor{red}{\sigma_k} \norm{f}_{1,2,\mu^2} \leq \sigma_k.
+	\label{eqn:paley_zygmund}
+	\mathbb{P}(Z > \lambda \mathbb{E}(Z^p)) \geq \left[(1 - \lambda^p) \frac{\mathbb{E}(Z^p)}{(\mathbb{E}(Z^q))^{p/q}}\right]^{\frac{q}{q - p}}
 	\end{equation}
-From Lemma \ref{lem: continuity_of_risk_functional}
-\begin{equation*}
-\lim_{n \to \infty} \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_n(z_k) = \int_{\Dset} f d\Pbb(x) - \int_{\Dset} f d\Qbb(x) = \gamma_{\Wset^{1,2}(\Dset,\mu^2)}.
-\end{equation*}
-and along with \eqref{eqn: limsup_discrete_energy}, this implies
-\begin{equation*}
-\lim_{n \to \infty} T_2(\lbf; G_{n,r}) \leq \gamma_{\Wset^{1,2}(\Dset,\mu^2)}(\Pbb,\Qbb).
-\end{equation*}
-
-Let $\Lbf^{\dagger}$ be the pseudoinverse of the Laplacian matrix $\Lbf$. We introduce $f_n^{\star} = \lbf^T \Lbf^{\dagger} \lbf$, which satisfies
-\begin{equation*}
-\abs{\int_{\Dset}f_n^{\star}d\Pbb_n - \int_{\Dset}f_n^{\star}d\Qbb_n} = T_2(\lbf; G_{n,r})
-\end{equation*}
-Note that
-\begin{equation*}
-\norm{\Bbf \fbf_n^{\star}}_2^2 \leq C_n \Longrightarrow \frac{1}{n^2 r_n^{d+2}} \norm{\Bbf \fbf_n^{\star}}_2^2  \leq \sigma_k < \infty
-\end{equation*}
-and so by the compactness property in Theorem \ref{thm: gamma_convergence_energy_functional}, every subsequence of $f_n^{\star}$ has a further subsequence which is $TL^2$-convergent.
-For simplicity, and without loss of generality, let us work along a subsequence which is convergent (and call it $f_n^{\star}$) to $f \in L^2(\mu)$.
-Then
-\begin{equation}
-\label{eqn: lower_bound_T2_1}
-\lim_{n \to \infty} \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_n^{\star}(z_k) = \abs{\int_{\Dset}f d\Pbb(x) - \int_{\Dset}f d\Qbb(x)}
-\end{equation}
-and by the liminf inequality of Theorem \ref{thm: gamma_convergence_energy_functional},
-\begin{equation*}
-\liminf_{n \to \infty} \frac{1}{n^2 r_n^{d+2}} \norm{\Bbf \fbf_n}_2^2 \geq \sigma_k \norm{f}_{1,2,\mu^2}
-\end{equation*}
-which implies $\norm{f}_{1,2,\mu^2} \leq 1$. This, along with \eqref{eqn: lower_bound_T2_1}, implies
-\begin{equation*}
-\lim_{n \to \infty} \frac{1}{n} \sum_{k = 1}^{n} \ell_k f_n^{\star}(z_k) \geq \gamma_{\Wset^{1,2}(\Dset,\mu^2)}(\Pbb,\Qbb)
-\end{equation*}
-
-\clearpage
-
-\bibliography{../graph_testing_bibliography}
-\bibliographystyle{plain}
+\end{lemma}
 

### 2019-12-10 notes/graph_testing_summary.tex
added high-level summary and project-level notes

--- 2019-12-10 notes/graph_testing.tex
@@
 
-\section{Proofs}
-
-\subsection{Proof of Theorem~\ref{thm:sobolev_testing_rate}}
-
-To prove Theorem~\ref{thm:sobolev_testing_rate}, we will show that when
-\begin{equation}
-\label{eqn:critical_radius}
-\epsilon^2 \geq c \cdot b \cdot (\log n)^{h(\beta,d)} \cdot n^{-4/(4 + d)}
-\end{equation}
-and we choose tuning parameters as specified in Theorem~\ref{thm:sobolev_testing_rate}, the test $\phi_{\mathrm{spec}}$ satisfies
-\begin{equation}
-\mathcal{R}_{\epsilon}(\phi_{\spec}; W^{1,2}(\mathcal{X};R)) = \Ebb_{f_0}(\phi_{\spec}) + \sup_{\substack{f \in W^{1,2}(\mathcal{X};R) \\ \norm{f}_{\Leb^2(\mathcal{X})} > \epsilon}} \Ebb_{f}(1 - \phi_{\spec}) \leq \left(\frac{1}{b^2} + \frac{3}{b}\right) + o(1).
-\end{equation}
-
-Our analysis will proceed according to the following steps.
-\begin{enumerate}
-	\item We upper bound the testing error of our eigenvector projection test statistic when $Y$ are viewed as random responses defined over the vertices of a fixed graph $G$.
-This upper bound will hold whenever certain functionals on the graph $G$ are themselves bounded.
-	\item We analyze the behavior of these functionals with respect to the random graph $G_{n,r}$, and bound them with high probability.
-	\item We condition on $X \subseteq \mathcal{E}$, where $\mathcal{E}$ is a high-probability set over which the relevant functionals on $G_{n,r}$ are bounded.
-We conclude that the upper bound on testing error derived in our first step holds whenever $X \subseteq \mathcal{E}$.
-\end{enumerate}
-
-\subsubsection{Step 1: Testing error on a fixed graph}
-
-Let $G = (V,E)$ be a graph over vertices $V = \set{v_1,\ldots,v_n}$, and let $\beta = (\beta_1,\ldots,\beta_n) \in \Reals^n$ be a signal over the vertices $V$.
-We observe responses $Y = (y_1,\ldots,y_n)$ according to the model
-\begin{equation*}
-y_i = \beta_i + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
-\end{equation*}
-Letting $L = VSV^T$ be the spectral decomposition of the Laplacian $L$ of $G$, our graph spectral test statistic is 
-\begin{equation*}
-T_{\spec} = \frac{1}{n}\sum_{k = 1}^{\kappa} \left(\sum_{i = 1}^{n} v_i y_i\right)^2
-\end{equation*}
-where $\kappa$ is a tuning parameter. The resulting test we will use is
-\begin{equation*}
-\phi_{\spec} = \1\{T_{\spec} \geq \frac{\kappa}{n} + t(b)\},~~\textrm{where}~ t(b) = b\sqrt{\frac{2\kappa}{n^2}}~~\textrm{for $b \geq 1$.}
-\end{equation*}
-
-Let $S_2(\beta;G)$ be a measure of smoothness the signal $\beta$ displays over the graph $G$, given by
-\begin{equation*}
-S_2(\beta;G) = \sum_{i = 1}^{n} \sum_{j = 1}^{n} (\beta_i - \beta_j)^2 \1((v_i,v_j) \in E).
-\end{equation*}
-In Lemma~, we upper bound the Type I and Type II error of the test $\phi_{\spec}$.
-Our bound on the Type II error will be stated as a function of $S_2(\beta;G)$ as well as the $\kappa$th eigenvalue $s_{\kappa}$.
-
-\begin{lemma}
-	\label{lem:fixed_graph_testing}
-	Let $1 \leq \kappa \leq n$ be an integer.
-	\begin{enumerate}
-		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}$ is upper bounded
-		\begin{equation}
-		\label{eqn:graph_spectral_type_I_error}
-		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{1}{b^2}.
-		\end{equation}
-		\item \textbf{Type II error:} For any $b$ and $\beta$ such that
-		\begin{equation}
-		\label{eqn:fixed_graph_testing_critical_radius}
-		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_2(\beta;G)}{ns_{\kappa}}
-		\end{equation}
-		the Type II error of $\phi_{\beta}$ is upper bounded,
-		\begin{equation}
-		\label{eqn:graph_spectral_type_II_error}
-		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{3}{b}.
-		\end{equation}
-	\end{enumerate}
-\end{lemma}
-
-The first term on the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius} grows with $\kappa$, while the second term shrinks.
-These can be thought of as the variance and squared bias terms, respectively, of the error.
-
-\paragraph{Proof of Lemma~\ref{lem:fixed_graph_testing}.}
-
-To prove Lemma~\ref{lem:fixed_graph_testing} we will first compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}$, and then use Chebyshev's inequality to show \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.
-
-\vspace{.2 in}
-
-\textit{Mean of $T_{\mathrm{spec}}$:} Using the notation $\dotp{v}{w} = \sum_{i = 1}^{n} v_iw_i$, we have
-\begin{align*}
-\Ebb(T_{\spec}) & = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{\beta}{v_k}^2 + \Ebb\bigl( \dotp{\varepsilon}{v_k}^2 + 2 \dotp{\varepsilon}{v_k} \dotp{\beta}{v_k}\bigr)\right) \\
-& = \frac{\kappa}{n} + \frac{1}{n}\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2.
-\end{align*}
-When $\beta = 0$, this equals $\kappa/n$. Otherwise, we have the following lower bound:
-\begin{align*}
-\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2 & = \norm{\beta}_2^2 - \sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \\
-& \geq \norm{\beta}_2^2 - \frac{1}{s_{\kappa}}\sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 s_k \\
-& \geq \norm{\beta}_2^2 - \frac{S_2(\beta;G)}{s_{\kappa}},
-\end{align*}
-and therefore $\Ebb(T_{\spec}) \geq \kappa/n + n^{-1}(\norm{\beta}_2^2 - S_2(\beta;G)/s_{\kappa})$.
-
-\vspace{.2 in}
-
-\textit{Variance of $T_{\mathrm{spec}}$:}
-We write $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvectors $v_1,\ldots,v_{\kappa}$ as columns. Consequently,
-\begin{align}
-\Var(T_{\spec}) & = \frac{1}{n^2} \Var(y^T V_{\kappa} V_{\kappa}^T y) \\
-& = \frac{1}{n^2} \Var((\beta + \varepsilon)^T V_{\kappa} V_{\kappa}^T (\beta + \varepsilon)) \\
-& = \frac{1}{n^2} \Var(2 \beta^T V_{\kappa} V_{\kappa}^T \varepsilon + \varepsilon^T V_{\kappa} V_{\kappa}^T \varepsilon) \\
-& \leq \frac{1}{n^2}(4 \beta^T V_{\kappa} V_{\kappa}^T \beta + 2\kappa)
-\end{align}
-where the last inequality follows from standard properties of the Gaussian distribution.
-We now move on to showing the desired inequalities \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.
-For simplicity, we introduce the notation
-\begin{equation*}
-t(b) = b\sqrt{\frac{2\kappa}{n^2}}, ~~ \Delta = \norm{\beta}_2^2 - \frac{S_2(\beta;G)}{s_{\kappa}}.
-\end{equation*}
-
-\vspace{.2 in}
-
-\textit{Proof of~\eqref{eqn:graph_spectral_type_I_error}:} By Chebyshev's inequality,
-\begin{align*}
-\Pbb_{\beta = 0}\bigl(T_{\spec} \geq \frac{\kappa}{n} + t(b)\bigr)
-& \leq \Pbb_{\beta = 0}\bigl(\abs{T_{\spec} - \frac{\kappa}{n}} \geq t(b)\bigr) \\
-& \leq \frac{\Var_{\beta = 0}(T_{\spec})}{t(b)^2} = \frac{1}{b^2}.
-\end{align*}
-
-\vspace{.2 in}
-
-\textit{Proof of~\eqref{eqn:graph_spectral_type_II_error}:} Assumption~\eqref{eqn:fixed_graph_testing_critical_radius} implies
-\begin{equation*}
-\Delta \geq \frac{\norm{\beta}_2^2}{n} - \frac{C_n^2}{nC^2} \geq 2 t(b).
-\end{equation*}
-Then another application of Chebyshev's inequality gives us
-\begin{align*}
-\Pbb_{\beta}\bigl(T_{\spec} \leq \frac{\kappa}{n} + t(b)\bigr) & = \Pbb_{\beta}\bigl(T_{\spec} - \Ebb_{\beta}(T_{\spec}) \leq t(b) - \Delta \bigr) \\
-& \leq \Pbb_{\beta}\bigl(\abs{T_{\spec} - \Ebb_{\beta}(T_{\spec})} \leq \Delta - t(b) \bigr) \tag{since $\Delta \geq t(b)$}	\\
-& \leq \frac{\Var_{\beta}(T_{\spec})}{(\Delta - t(b))^2} \\
-& \leq 4\frac{\Var_{\beta}(T_{\spec})}{\Delta^2} \tag{since $\Delta \geq 2t(b)$} \\
-& \leq 4\frac{2\kappa/n^2 + 4\beta^T V_{\kappa} V_{\kappa}^T \beta /n}{\Delta^2}.
-\end{align*}
-We handle each summand in the numerator separately. For the first term, since $\Delta \geq 2 t(b)$, we have
-\begin{equation}
-\label{eqn:spectral_type_II_error_pf1}
-\frac{2\kappa}{n^2\Delta^2} \leq \frac{1}{2b^2}.
-\end{equation}
-
-For the second term, noting that $\Delta = \beta^T V_{\kappa} V_{\kappa}^T \beta$, we have
-\begin{align}
-\frac{\beta^T V_{\kappa} V_{\kappa}^T \beta/n}{\Delta^2} & \leq \frac{1}{n\Delta} \nonumber \\
-& \leq \frac{1}{2nt(b)} \nonumber \\
-& = \frac{1}{2b\sqrt{2\kappa}}, \label{eqn:spectral_type_II_error_pf2}
-\end{align}
-and combining~\eqref{eqn:spectral_type_II_error_pf1} and~\eqref{eqn:spectral_type_II_error_pf2} yields~\eqref{eqn:graph_spectral_type_II_error}.
-
-\subsubsection{Step 2: Bounding neighborhood graph functionals}
-
-To make use of Lemma~\ref{lem:fixed_graph_testing} we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$.
-In particular, we will show that for some constants $c_1,c_2,c_3$, the following statements:
-\begin{enumerate}[label=(E\arabic*)]
-	\item 
-	\label{event:discrete_sobolev_norm}
-	\textbf{Graph Sobolev norm:} When $\beta = (f(x_1),\ldots,f(x_n))$ for $f \in W^{1,2}(\mathcal{X};R)$:
-	\begin{equation}
-	\label{eqn:continuous_to_discrete_sobolev_norm}
-	S_2(\beta;G_{n,r}) \leq c_1 \cdot n^2 r^{d + 2} 
-	\end{equation}
-	\item 
-	\label{event:eigenvalue_tail_decay}
-	\textbf{Eigenvalue tail bound:} If $r = (\log^{a}n)\cdot(\log n/n)^{1/d}$ for any $a > 0$, then for all $\kappa \in [n]$:
-	\begin{equation}
-	\label{eqn:eigenvalue_tail_bound}
-	s_{\kappa} \geq c_2 \cdot \frac{\kappa^{2/d}}{n^{2/d}}
-	\end{equation}
-	\item 
-	\label{event:l2_norm}
-	\textbf{Empirical norm of $f$:} When $f \in W^{1,2}(\mathcal{X};R)$ and one of (a) $d = 1,2$, or (b) $d = 3$ and $\norm{f}_{\Leb^2}^2 \geq n^{-2/(d - 2)}$, then
-	\begin{equation}
-	\label{eqn:l2_to_empirical_norm}
-	\frac{1}{n}\sum_{i = 1}^{n}\beta_i^2 = \norm{f}_n^2 \geq c_3 \cdot \norm{f}_{\Leb^2}^2
-	\end{equation}
-\end{enumerate}
-
-each hold with probability $1 - o(1)$ as $n \to \infty$.
-
-\paragraph{Proof of~\eqref{eqn:continuous_to_discrete_sobolev_norm}:}
-
-Observe that
-\begin{align*}
-\frac{1}{n^2}\Ebb(\beta^T L \beta) = \Ebb\left(\frac{1}{n^2} \sum_{i,j = 1}^{n} (\beta_i - \beta_j)^2 A_{ij}\right) & = \frac{(n - 1)}{n} \int_{\Xset} \int_{\Xset} (f(x) - f(y))^2K_r(x,y) \,dP(y) \,dP(x) \\
-& \leq p_{\max}^2 \frac{(n - 1)}{n} \int_{\Xset} \int_{\X} (f(x) - f(y))^2K_r(x,y) \,dy \,dx.
-\end{align*}
-We will show that for any $f \in \mathcal{W}^{1,2}(\X;R)$
-\begin{equation}
-\label{eqn:nonlocal_continuous_sobolev_norm}
-\int_{\Xset} \int_{\Xset} (f(x) - f(y))^2K_r(x,y) \,dy \,dx \leq c r^{d + 2} \norm{f}_{W^{1,2}(\mathcal{X})}
-\end{equation}
-whence the desired result~\eqref{eqn:continuous_to_discrete_sobolev_norm} follows by Markov's inequality and the assumption $f \in W^{1,2}(\mathcal{X};R)$.
-
-We begin by dealing with complications due to smoothing at the boundary of $\mathcal{X} = [0,1]^d$.
-Let $U$ be any bounded open set such that $\Xset \subset U$ is compactly contained in $U$.
-Note that as $\partial \Xset$ is $C^1$, by Theorem \ref{thm:evans_extension} there exists $g \in W^{1,2}(\Reals^d)$ such that
-\begin{enumerate}[(i)]
-	\item
-	\label{eqn:evans_extension_1}
-	$g = f$, $P$-almost-everywhere in $\Xset$
-	\item 
-	$g$ has support within $V$, and  
-	\item 
-	\label{eqn:sobolev_1_bound_2}
-	$\norm{g}_{W^{1,2}(\Rd)} \leq c \norm{f}_{W^{1,2}(\Xset)}$ for a constant $c$ which depends only on $\Xset$.
-\end{enumerate}
-As a result of \ref{eqn:evans_extension_1},
-\begin{equation}
-\label{eqn:sobolev_1_bound_1}
-\int_{\Xset} \int_{\Xset} (f(x) - f(y))^2K_r(x,y) \,dy \,dx \leq \int_{\Rd} \int_{\Rd} (g(x) - g(y))^2K_r(x,y) \,dy \,dx.
-\end{equation}
-
-Next we smooth $g$, so that we may work with ordinary partial derivatives.
-We let $\eta \in C^{\infty}(\Rd)$ be given by
-\begin{equation*}
-\eta(x) :=
-\begin{cases}
-C \exp \left\{\frac{1}{\norm{x}^2 - 1}\right\}& \quad \textrm{if $\norm{x}_2 \leq 1$} \\
-0 & \quad \textrm{if $\norm{x}_2 \geq 1$}
-\end{cases}
-\end{equation*}
-where the normalizing constant $C > 0$ is chosen so that $\int_{\Rd} \eta dx = 1$. Let $\eta_r(x) := (1/r^d) \eta(x/r)$. Then, the mollification of $g$ by $\eta_r$ is given by
-\begin{align*}
-g^r & := g \ast \eta_r \\
-& = \int_{\Rd} \eta_r(x - y) g(y) dy
-\end{align*}
-(Refer to \textcolor{red}{Evans}, Appendix C, Theorem 7 for a proof that $g^r \in C^{\infty}(\Rd)$.)
-Adding and subtracting within \eqref{eqn:sobolev_1_bound_1}, we have
-\begin{align}
-\int_{\Rd} & \int_{\Rd} (g(x) - g(y))^2K_r(x,y) \,dy \,dx \\
-& \leq 3 \int_{\Rd} \int_{\Rd} \bigl((g(x) - g^r(x))^2 + (g^r(x) - g^r(y))^2 + (g^r(y) - g(y))^2\bigr)K_r(x,y) \,dy \,dx \nonumber \\
-& = 6 \int_{\Rd} \int_{\Rd} \bigl((g(y) - g^r(y))^2 K_r(x,y) \,dy \,dx + 2 \int_{\Rd} \int_{\Rd} \bigl((g^r(x) - g^r(y))^2 K_r(x,y) \,dy \,dx \label{eqn:sobolev_1_bound_5}
-\end{align}
-We deal with each summand individually, beginning with the first one. Let $\nu_d = \pi^{d/2}/\Gamma(d/2 + 1)$ denote the volume of a ball with unit radius in $\Rd$. Then, we have
-\begin{align}
-\int_{\Rd} \int_{\Rd} \bigl((g(y) - g^r(y))^2 K_r(x,y) \,dy \,dx & = \int_{\Rd} \int_{B(x,r)} \bigl((g(y) - g^r(y))^2\,dy \,dx \nonumber \\
-& = \int_{\Rd} \int_{B(y,r)} \bigl((g(y) - g^r(y))^2\,dx \,dy \tag{by Tonelli's Theorem} \nonumber \\
-& = \nu_d r^d \int_{\Rd}\bigl((g(y) - g^r(y))^2 \,dy \\
-& \leq \nu_d r^{d + 2} \int_{\Rd} \norm{\nabla g(y)}^2 \,dy \label{eqn:sobolev_1_bound_4}
-\end{align}
-where the last line follows from Lemma \ref{lem:poincare_mollify}.
-
-We now turn our attention to the second summand. Note that as $g^r \in C^{\infty}(\Rd)$, we may apply Theorem \ref{thm:taylor_expansion} and obtain
-\begin{align*}
-(g^r(x) - g^r(y))^2 & = \left(\int_{0}^{1} \nabla g^r(x + t(y - x)) \cdot (y - x) \,dt\right)^2 \\
-& \leq \int_{0}^{1} \bigl(\nabla g^r(x + t(y - x)) \cdot (y - x)\bigr)^2 \,dt  \tag{Jensen's inequality} \\
-& \leq \norm{y - x}^2 \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 \,dt. \tag{Cauchy-Schwarz inequality}
-\end{align*}
-As a result, we have
-\begin{align*}
-\int_{\Rd} \int_{\Rd} \bigl((g^r(x) - g^r(y))^2 K_r(x,y) \,dy) \,dx & \leq \int_{\Rd} \int_{\Rd} \norm{y - x}^2  \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 K_r(x,y) \,dt \,dy \,dx \\
-& \leq r^{2} \int_{\Rd} \int_{\Rd} \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 K_r(x,y) \,dt \,dy \,dx \\
-& = r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + t(y - x))}^2 K_r(x,y) \,dx \,dt \,dy \\
-& = r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + tz)}^2 K_r(z) \,dx \,dt \,dz \tag{$z = y - x$}
-\end{align*}
-
-where we write $K_r(z) = \mathbf{1}(\norm{z} \leq r)$ in an abuse of notation. Next, we note that
-\begin{equation*}
-\int_{\Rd} \norm{\nabla g^r(x + tz)}^2 \,dx = \int_{\Rd} \norm{\nabla g^r(x)}^2 \,dx \leq \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
-\end{equation*}
-with the inequality following from Lemma \ref{lem:gradient_mollify_commute}. Therefore,
-\begin{align}
-r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + tz)}^2 K_r(z) \,dx \,dt \,dz & \leq r^{2} \int_{\Rd} K_r(z) \int_{0}^{1} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx  \,dt \,dz \nonumber \\
-& = \nu_d r^{2 + d} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx. \label{eqn:sobolev_1_bound_3}
-\end{align}
-By \eqref{eqn:sobolev_1_bound_5}, \eqref{eqn:sobolev_1_bound_4} and \eqref{eqn:sobolev_1_bound_3}, we have that 
-\begin{equation*}
-\int_{\Rd} \int_{\Rd} (g(x) - g(y))^2 K_r(x,y) \,dx \,dy \leq 9 \nu_d r^{d + 2} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
-\end{equation*}
-Then by \ref{eqn:sobolev_1_bound_2}, $\int_{\Rd} \norm{\nabla g(x)}^2 \,dx \leq c \int_{\Xset} \norm{\nabla f(x)}^2 \,dx$, and the desired result of \eqref{eqn:sobolev_1_bound_1} follows.
-
-\paragraph{Proof of~\eqref{eqn:eigenvalue_tail_bound}:}
-
-For ease of notation we write $t = n^{1/d}$, and assume without loss of generality that $t$ is whole number. Let $\overline{G} = (\overline{V},\overline{E})$ be the lattice over $[0,1]^d$, formally
-\begin{align*}
-\overline{V} & = \bigl\{\overline{x}_k := \frac{1}{n}(k_1,\ldots,k_d): k \in [t]^d\bigr\} \\
-\overline{E} & = \bigl\{(\overline{x}_k, \overline{x}_{\ell}): \norm{k - \ell}_1 = 1\bigr\}.
-\end{align*}
-
-By Theorem 1.1 of \textcolor{red}{Garcia Trillos and Slepcev}, with probability at least $1 - n^{-1}$ there exists a bijection $\pi: \overline{V} \to X$ such that
-\begin{equation*}
-\max_{k \in [t]^d} \abs{\overline{x}_k - \pi(\overline{x}_k)} \leq c \left(\frac{\log n}{n}\right)^{1/d}
-\end{equation*}
-Since our choice $r = \log^a(n) \cdot (\log n/n)^{1/d}$ implies $r = \omega(\left(\frac{\log n}{n}\right)^{1/d})$, for $n$ sufficiently large with probability $1 - n^{-1} = 1 - o(1)$ we have
-\begin{equation}
-\label{eqn:eigenvalue_tail_bound_pf1}
-(\overline{x}_k,\overline{x}_{\ell}) \in \overline{E} \Longrightarrow (\pi(\overline{x}_k),\pi(\overline{x}_{\ell})) \in E.
-\end{equation}
-Let $\overline{L}$ be the Laplacian of $\overline{G}$, with eigenvalues $\lambda_1 < \cdots \leq \lambda_n$. If \eqref{eqn:eigenvalue_tail_bound_pf1} holds, then for all $x \in \Reals^n$,
-\begin{equation*}
-x^T \overline{L} x \geq x^T L x \Longrightarrow \lambda_j \leq s_j~~ \textrm{for all}~ j \in [n].
-\end{equation*}
-The eigenvalues of the $d$-dimensional lattice are explicitly known.
-In particular when $j = h^d$ for some whole number $h$, we have $s_j \geq \lambda_j = 4 d \sin^2(\pi h/2t) \geq \pi^2 h^2/ (4 t^2)$, with a similar inequality (for possible different constants) holding when $j^{1/d}$ is not integral; this shows~\eqref{eqn:eigenvalue_tail_bound}.
-
-\paragraph{Proof of~\eqref{eqn:l2_to_empirical_norm}:}
-
-To prove~\eqref{eqn:l2_to_empirical_norm} we will upper bound
-\begin{equation*}
-\mathbb{E}\biggl[\Bigl(\frac{1}{n}\sum_{i = 1}^{n}f^2(x_i)\Bigr)^2\biggr],
-\end{equation*}
-and then apply the Paley-Zygmund inequality (Lemma~\ref{lem:paley_zygmund}). We have
-\begin{align*}
-\mathbb{E}\biggl[\Bigl(\frac{1}{n}\sum_{i = 1}^{n}f^2(x_i)\Bigr)^2\biggr] & = \frac{(n-1)}{n}\mathbb{E}(f^2(x_1)) + \frac{\mathbb{E}[f^4(x_1)]}{n} \\
-& \leq p_{\max}^2 \left(\norm{f}_{\Leb^2}^2 + \frac{\norm{f}_{\Leb^4}^4}{n}\right) \\
-& \overset{(i)}{\leq} p_{\max}^2 \left(\norm{f}_{\Leb^2}^2 + c\frac{\norm{f}_{\Leb^2}^{(4 - d)}}{n}\right) \\
-& \overset{(ii)}{\leq}  p_{\max}^2 \left(\norm{f}_{\Leb^2}^2 + c\norm{f}_{\Leb^2}^2\right).
-\end{align*}
-Here $(i)$ follows from two facts, the first being the embedding $W^{1,2}(\mathcal{X};R) \subseteq \Leb^q(\mathcal{X};R')$ for $q = 2d/(d - 2)$ when $d > 2$ and $q = \infty$ when $d \leq 2$ and some constant $R'$, and the second being a standard result on interpolation of $\Leb^p$ spaces, namely that since $f \in \Leb^2(\mathcal{X}) \cap \Leb^q(\mathcal{X})$, then
-\begin{equation*}
-\norm{f}_{\Leb^4} \leq \norm{f}_{\Leb^2}^{\lambda} \cdot \norm{f}_{\Leb^q}^{1 - \lambda},~~\textrm{where}~ \frac{1}{4} = \frac{\lambda}{2} + \frac{1 - \lambda}{q}.
-\end{equation*} 
-Then $(ii)$ follows from our assumptions on $\norm{f}_{\Leb^2}^2$: either 
-\begin{enumerate}[(a)]
-	\item $d \leq 2$ and $\norm{f}_{\Leb^2}^2 \geq 1/n$ so $\norm{f}_{\Leb^2}^{(4 - d)} \leq \max(R^{4 - d}/n, \norm{f}_{\Leb^2}^2) \leq c\norm{f}_{\Leb^2}^2$, or
-	\item $d = 3$ and $\norm{f}_{\Leb^2}^2 \geq n^{-2/(d - 2)}$.
-\end{enumerate} 
-Applying Lemma~\ref{lem:paley_zygmund} to  $\norm{f}_n^2$ yields~\eqref{eqn:l2_to_empirical_norm}.
-
-\subsubsection{Step 3: Conclusion}
-
-We note that for all possible values of $X \in \Xset^n$, under the null hypothesis $f = f_0 = 0$ and therefore $\beta = (f(x_1),\ldots,f(x_n)) = 0$ as well.
-Therefore by~\eqref{eqn:graph_spectral_type_I_error}, we have the following bound on Type I error:
-\begin{equation}
-\Ebb_{f_0}(\phi_{\mathrm{spec}}) = \mathbb{E}(\mathbb{E}_{\beta = 0}(\phi_{\spec}) | X) \leq \frac{1}{b^2}.
-\end{equation}
-
-Now, we bound Type II error under the assumption $f \in W^{1,2}(\mathcal{X};R)$ and 
-\begin{equation}
-\label{eqn:critical_radius_1}
-\norm{f}_{\Leb^2}^2 \geq \epsilon^2 = c \cdot b \cdot (\log n)^{h(\beta,d)} \cdot n^{-4/(4 + d)}.
-\end{equation}
-We verify that the conditions required to apply Step 2 are satisfied.
-In particular, we choose $r = (\log^a n) \cdot (\log n/n)^{1/d}$, and observe that~\eqref{eqn:critical_radius_1} implies $\norm{f}_{\Leb^2}^2 \geq n^{-1}$ when $d \leq 2$ and $\norm{f}_{\Leb^2}^2 \geq n^{-2/(d - 2)}$ when $2 < d < 4$.
-We may therefore apply our conclusions in Step 2; namely, that for every possible choice of $f$ there exists a good set $\mathcal{E}_f \subseteq \Xset^n$ with $\Pbb(\mathcal{E}_f) \geq 1 - o(1)$ such that each of \eqref{eqn:continuous_to_discrete_sobolev_norm}, \eqref{eqn:eigenvalue_tail_bound}, and \eqref{eqn:l2_to_empirical_norm} hold for all $X \subseteq \mathcal{E}_f$.
-Choosing $\kappa = n^{2d/(4 + d)}$ to balance the squared bias and variance terms on the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius}, we have that for all $X \subseteq \mathcal{E}_f$
-\begin{align*}
-2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_2(\beta;G)}{ns_{\kappa}} & \leq 2bn^{-4/(4+d)} + \frac{n r^{d + 2}}{s_{\kappa}} \tag{by \eqref{eqn:continuous_to_discrete_sobolev_norm}} \\
-& \leq 2bn^{-4/(4+d)} + c\cdot\frac{n r^{d + 2}}{n^{4/(4+d)}} \tag{by \eqref{eqn:eigenvalue_tail_bound}} \\
-& \leq 2bn^{-4/(4+d)} + c\cdot(\log^{a(d + 2) + (1 + 2/d)}n)\cdot n^{-4/(4+d)} \\
-& \leq \frac{1}{10}\norm{f}_{\Leb^2} \tag{for a suitably large choice of $c$ in \eqref{eqn:critical_radius_1}}\\
-& \leq \frac{1}{n}\sum_{i = 1} \beta_i^2. \tag{by \eqref{eqn:l2_to_empirical_norm}}
-\end{align*}
-We conclude that for all $X \subseteq \mathcal{E}_f$, the inequality \eqref{eqn:fixed_graph_testing_critical_radius} is satisfied with respect to $\beta = (f(x_1),\ldots,f(x_n))$ and $G = G_{n,r}$.
-As a result the worst-case Type II error is bounded
-\begin{equation*}
-\sup_{\substack{f \in W^{1,2}(\mathcal{X;R}) \\ \norm{f}_{\Leb^2} \geq \epsilon}}\mathbb{E}_{f}(1 - \phi_{\spec}) \leq \sup_{\substack{f \in W^{1,2}(\mathcal{X;R}) \\ \norm{f}_{\Leb^2} \geq \epsilon}} \mathbb{E}\bigl[\mathbb{E}_{\beta}(1 - \phi_{\spec}|X \in \mathcal{E}_f)\bigr] + o(1) \leq \frac{3}{b} + o(1).
-\end{equation*}
-
-\subsection{Proof of Proposition~\ref{prop:L4_testing_rate}}
-
-Let $T_{\mathrm{mean}} = \frac{1}{n}\sum_{i = 1}^{n} y_i^2$. The expectation of $T_{\mathrm{mean}}$ is
-\begin{equation*}
-\Ebb(T_{\mathrm{mean}}) = \mathbb{E}(f^2(x_1)) + 1,
-\end{equation*}
-and the variance can be upper bounded
-\begin{equation*}
-\Var(T_{\mathrm{mean}}) \leq \frac{1}{n}(3 + p_{\max} \norm{f}_{\Leb^4}^4 + p_{\max}\norm{f}_{\Leb^2}^2) = \frac{c}{n}.
-\end{equation*}
-Since $\mathbb{E}(f^2(x_1)) \geq p_{\min} \epsilon^2$, when $\epsilon^2 \gtrsim b\cdot n^{-1/2}$ we can apply Chebyshev's inequality to obtain the claimed result.
-\section{Auxiliary Results}
-
-\begin{theorem}[\textcolor{red}{Evans} Chapter 5.4, Theorem 1]
-	\label{thm:evans_extension}
-	Assume $U$ is bounded and $\partial U$ is $C^1$.
-Select a bounded open set $V$ such that $U \subset \subset V$ ($U$ is compactly contained in $V$).
-Then there exists a bounded linear operator $E: W^{1,2}(U) \to W^{1,2}(\Rd)$ such that for each $u \in W^{1,2}(U)$:
-	\begin{enumerate}
-		\item $Eu = u$ a.e. in $U$,
-		\item $Eu$ has support within $V$, and 
-		\item 
-		\begin{equation*}
-		\norm{Eu}_{W^{1,2}(\Rd)} \leq C \norm{u}_{W^{1,2}(\Rd)}
-		\end{equation*}
-		the constant $C$ depending only on $U$ and $V$.
-	\end{enumerate}
-\end{theorem}
-
-For $u \in W^{1,2}(\Rd)$ and $x \in \Rd$, write $\nabla u(x) = (D^{e_1}(x),\ldots,D^{e_d}(x)$, for the gradient of $u$.
-
-\begin{theorem}[\textcolor{red}{Evans} Chapter 5.8.1, Theorem 2]
-	\label{thm:evans_poincare}
-	There exists a constant $C$, depending only on $d$, such that
-	\begin{equation*}
-	\norm{u - u^r}_{L^2(B(x,r))} \leq Cr\norm{\nabla u}_{L^2(B(x,r))}
-	\end{equation*}
-\end{theorem}
-
-\begin{theorem}[Taylor expansion.]
-	\label{thm:taylor_expansion}
-	For any function $u \in C^1$, and any $x,y \in \Rd$, 
-	\begin{equation*}
-	u(y) - u(x) = \int_{0}^{1} \nabla(u(x + t(y - x))) \cdot (y - x) \,dt
-	\end{equation*}
-\end{theorem}
-
-\begin{lemma}
-	\label{lem:poincare_mollify}
-	For any function $g \in W^{1,2}(\Rd)$ compactly supported in a bounded open set $V \subset \Rd$, we have
-	\begin{equation}
-	\label{eqn:poincare_mollify}
-	\int_{\Rd} (g(x) - g^r(x))^2 \,dx \leq r^2 \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
-	\end{equation}
-\end{lemma}
-\begin{proof}
-	This Lemma is essentially a reproduction of part of the proof of the Rellich-Kondrachov Compactness Theorem from \textcolor{red}{Evans}.
-Note that it is sufficient to prove in the case when $g$ is smooth.
-To see this, for the moment assume \eqref{eqn:poincare_mollify} holds for all $u \in C^{\infty}(V)$, and let $g \in W^{1,2}(V)$.
-By Theorem \ref{thm:local_approx_smooth_functions}, we may take a sequence $g_m \in C^{\infty}(V)$ such that
-	\begin{equation*}
-	\norm{g - g_m}_{L^2(V)} \to 0, \quad \textrm{and} \quad \norm{\nabla g - \nabla g_m}_{L^2(V)} \to 0,
-	\end{equation*}
-	Then we have
-	\begin{align*}
-	\int_{\Rd} (g(x) - g^r(x))^2 \,dx & \leq \int_{\Rd} (g(x) - g_m(x))^2 \,dx + \int_{\Rd}(g_m(x) - g_m^r(x))^2 \,dx + \int_{\Rd}(g_m^r(x) - g^r(x))^2 \,dx \\
-	& \leq \int_{\Rd} (g(x) - g_m(x))^2 \,dx + r^2 \int_{\Rd} \norm{\nabla g_m(x)}^2 \,dx + \int_{\Rd}(g_m^r(x) - g^r(x))^2 \,dx
-	\end{align*}
-	and taking the limit as $m$ goes to infinity, the right hand side converges to $\int_{\Rd} \norm{\nabla g(x)}^2 \,dx$.
-
-	It remains to show \eqref{eqn:poincare_mollify} in the case where $g$ is smooth. In this case, we have
-	\begin{align*}
-	g^{r}(x) -  g(x) & = \frac{1}{r^d} \int_{B(x,r)} \kappa \left(\frac{x - z}{r}\right)\bigl(g(z) - g(x)\bigr) \,dz \\
-	& = \int_{B(0,1)} \kappa(y) \bigl(g(x - ry) - g(x)\bigr) \,dy \\
-	& = \int_{B(0,1)} \kappa(y) \int_{0}^{1} \frac{d}{dt} \bigl(g(x - try) \bigr) \,dt \,dy \\
-	& = -r \int_{B(0,1)} \kappa(y) \int_{0}^{1} \bigl(\nabla g (x - try) \bigr) \cdot y \,dt \,dy.
-	\end{align*} 
-	Therefore, by Jensen's and Cauchy-Schwarz inequalities, we have
-	\begin{align*}
-	\int_{\Rd} (g(x) - g^r(x))^2 \,dx & \leq r^2 \int_{\Rd} \int_{B(0,1)} \kappa(y) \int_{0}^{1} \norm{\nabla g (x - try)}^2 \norm{y}^2 \,dt \,dy \,dx \\
-	& \leq r^2 \int_{B(0,1)} \kappa(y) \int_{0}^{1} \int_{\Rd} \norm{\nabla g (x - try)}^2 \,dx \,dt \,dy \\
-	& = r^2 \int_{\Rd} \norm{\nabla g(z)}^2 \,dz
-	\end{align*}
-\end{proof}
-
-The following theorem is Theorem 1 in Section 5.3 of \textcolor{red}{Evans}.
-\begin{theorem}[Local approximation by smooth functions.]
-	\label{thm:local_approx_smooth_functions}
-	Assume $U$ is bounded, and $u \in W^{k,p}(U)$ for some $1 \leq p < \infty$. Then there exists functions $u_m \in C^{\infty}(U) \cap W^{k,p}(U)$ such that
-	\begin{equation*}
-	\norm{u_m - u}_{W^{k,p}(U)} \overset{m}{\to} 0.
-	\end{equation*}
-\end{theorem}
-
-\begin{lemma}
-	\label{lem:gradient_mollify_commute}
-	For any $u \in W^{1,2}(\Rd)$, we have
-	\begin{equation*}
-	\int_{\Reals^d} \norm{\nabla u^r(x)}_2^2 \,dx \leq \int_{\Reals^d} \norm{\nabla u(x)}_2^2 \,dx
-	\end{equation*}
-\end{lemma}
-\begin{proof}
-	Observe that $\norm{\nabla u(x)}_2^2 = \sum_{j = 1}^{d} (D^{e_j}u(x))^2 $. Therefore,
-	\begin{align*}
-	\int_{\Rd} \norm{\nabla u^r(x)}_2^2 \,dP(x) & = \sum_{j = 1}^{d} \int_{\Rd} (D^{e_j}u^r(x))^2 \,dx \\
-	& = \sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u)^r(x))^2 \,dx
-	\end{align*}
-	where the second equality follows from equation (1) in Section 5.3 of \textcolor{red}{Evans}. Then, for any $v \in L^2(\Rd)$, we have that
-	\begin{align*}
-	\abs{v^r(x)} & = \int_{\Rd} \kappa_r^{1/2}(x - y) \kappa_r^{1/2}(x - y) v(y) \,dy \\
-	& \leq \left(\int_{\Rd} \kappa_r^(x - y) \,dy\right)^{1/2} \left(\int_{\Rd} \kappa_r(x - y) v^2(y) \right)^{1/2} \\
-	& = \left(\int_{\Rd} \kappa_r(x - y) v^2(y) \right)^{1/2}
-	\end{align*}
-	and therefore
-	\begin{align*}
-	\int_{\Rd} (v^r(x))^2 \,dx & \leq \int_{\Rd} \int_{\Rd } \kappa_r(x - y) v^2(y) \,dy \,dx \\
-	& = \int_{\Rd} \int_{\Rd } \kappa_r(x - y) v^2(y) \,dy \,dx \\
-	& = \int_{\Rd} v^2(y) \,dy
-	\end{align*}
-
-	Applying this to $D^{e_j}u \in L^2(\Rd)$, we have that
-	\begin{equation*}
-	\sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u)^r(x))^2 \,dx \leq \sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u(x))^2 \,dx = \int_{\Rd} \norm{\nabla u(x)}^2 \,dx.
-	\end{equation*}
-\end{proof}
-
-\begin{lemma}
-	\label{lem:paley_zygmund}
-	Let $0 < p < q$, and let $Z$ be a non-negative random variable such that $\mathbb{E}(Z^q) < \infty$. Then, for all $0 \leq \lambda \leq 1$,
-	\begin{equation}
-	\label{eqn:paley_zygmund}
-	\mathbb{P}(Z > \lambda \mathbb{E}(Z^p)) \geq \left[(1 - \lambda^p) \frac{\mathbb{E}(Z^p)}{(\mathbb{E}(Z^q))^{p/q}}\right]^{\frac{q}{q - p}}
-	\end{equation}
-\end{lemma}
-
 \end{document}

### 2019-12-13 notes/graph_testing.tex
updated proofs

@@
 \newcommand{\thetast}{\theta^{\star}}
+\newcommand{\betap}{\beta^{(p)}}
+\newcommand{\betaq}{\beta^{(q)}}
+\newcommand{\vardeltapq}{\varDelta^{(p,q)}}
 
@@
 \newcommand{\wt}[1]{\widetilde{#1}}
+\newcommand{\ol}[1]{\overline{#1}}
 \newcommand{\spec}{\mathrm{spec}}
@@
 	Let $b \geq 1$ and $a > 0$ be fixed constants, and let $d = 1$.
-Suppose that $P$ and $Q$ are absolutely continuous probability measures over $\mathcal{X} = [0,1]$ with density functions $p$ and $q$ bounded above and below by constants, i.e
-	\begin{equation*}
-	0 < p_{\min},q_{\min} < p(x),q(x) < p_{\max},q_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
+Suppose that $\mu = (P + Q)/2$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]$ with density functions $\rho$ bounded above and below by constants, i.e
+	\begin{equation*}
+	0 < \rho_{\min} < \rho(x) < \rho_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
 	\end{equation*}
@@
 \label{eqn:critical_radius}
-\epsilon^2 \geq c \cdot b \cdot (\log n)^{h(\beta,d)} \cdot n^{-4/(4 + d)}
+\epsilon^2 \geq c \cdot b \cdot (\log n)^{h(a,d)} \cdot n^{-4/(4 + d)}
 \end{equation}
@@
 		\end{equation}
-		the Type II error of $\phi_{\beta}$ is upper bounded,
+		the Type II error of $\phi_{\spec}$ is upper bounded,
 		\begin{equation}
@@
 \begin{equation*}
-t(b) = b\sqrt{\frac{2\kappa}{n^2}}, ~~ \Delta = \norm{\beta}_2^2 - \frac{S_2(\beta;G)}{s_{\kappa}}.
+\Delta = \norm{\beta}_2^2 - \frac{S_2(\beta;G)}{s_{\kappa}}.
 \end{equation*}
@@
 
-For ease of notation we write $t = n^{1/d}$, and assume without loss of generality that $t$ is whole number. Let $\overline{G} = (\overline{V},\overline{E})$ be the lattice over $[0,1]^d$, formally
-\begin{align*}
-\overline{V} & = \bigl\{\overline{x}_k := \frac{1}{n}(k_1,\ldots,k_d): k \in [t]^d\bigr\} \\
+For ease of notation we write $t = n^{1/d}$, and assume without loss of generality that $t$ is whole number. Let $\overline{G} = (\overline{X},\overline{E})$ be the lattice over $[0,1]^d$, formally
+\begin{align*}
+\overline{X} & = \bigl\{\overline{x}_k := \frac{1}{t}(k_1,\ldots,k_d): k \in [t]^d\bigr\} \\
 \overline{E} & = \bigl\{(\overline{x}_k, \overline{x}_{\ell}): \norm{k - \ell}_1 = 1\bigr\}.
@@
 By Theorem 1.1 of \textcolor{red}{Garcia Trillos and Slepcev}, with probability at least $1 - n^{-1}$ there exists a bijection $\pi: \overline{V} \to X$ such that
-\begin{equation*}
+\begin{equation}
+\label{eqn:transport_distance}
 \max_{k \in [t]^d} \abs{\overline{x}_k - \pi(\overline{x}_k)} \leq c \left(\frac{\log n}{n}\right)^{1/d}
-\end{equation*}
+\end{equation}
 Since our choice $r = \log^a(n) \cdot (\log n/n)^{1/d}$ implies $r = \omega(\left(\frac{\log n}{n}\right)^{1/d})$, for $n$ sufficiently large with probability $1 - n^{-1} = 1 - o(1)$ we have
@@
 \end{equation}
-Let $\overline{L}$ be the Laplacian of $\overline{G}$, with eigenvalues $\lambda_1 < \cdots \leq \lambda_n$. If \eqref{eqn:eigenvalue_tail_bound_pf1} holds, then for all $x \in \Reals^n$,
-\begin{equation*}
-x^T \overline{L} x \geq x^T L x \Longrightarrow \lambda_j \leq s_j~~ \textrm{for all}~ j \in [n].
+Let $\overline{L}$ be the Laplacian of $\overline{G}$, with eigenvalues $\overline{s}_1 < \cdots \leq \overline{s}_n$. If \eqref{eqn:eigenvalue_tail_bound_pf1} holds, then for all $x \in \Reals^n$,
+\begin{equation*}
+x^T \overline{L} x \leq x^T L x \Longrightarrow \overline{s}_j \leq s_j~~ \textrm{for all}~ j \in [n].
 \end{equation*}
@@
 Since $\mathbb{E}(f^2(x_1)) \geq p_{\min} \epsilon^2$, when $\epsilon^2 \gtrsim b\cdot n^{-1/2}$ we can apply Chebyshev's inequality to obtain the claimed result.
+
+\subsection{Proof of Theorem~\ref{thm:twosample_sobolev_testing_rate}}
+
+As mentioned previously, to prove Theorem~\ref{thm:twosample_sobolev_testing_rate}, we relate the two sample model to the following regression model: we observe
+$X = \{x_1,\ldots,x_n\} \sim \mu$ (where we recall $\mu = (P + Q)/2$), and associated labels
+\begin{equation*}
+a_i = 
+\begin{cases}
+1, & \textrm{with probability}~ \frac{p(x_i)}{p(x_i) + q(x_i)} \\
+-1, & \textrm{with probability}~ \frac{q(x_i)}{p(x_i) + q(x_i)}
+\end{cases}
+\end{equation*}
+where $a_i$ is conditionally independent of $x_j,a_j$ given $x_i$.
+
+\textcolor{red}{TODO:} Complete the above argument.
+
+Our analysis will trace a similar path to the proof of Theorem~\ref{thm:sobolev_testing_rate}, proceeding according to the following steps:
+\begin{enumerate}
+	\item We upper bound the testing error of our eigenvector projection test statistic when $a$ are viewed as random responses defined over the vertices of a fixed graph $G$.
+This upper bound will hold whenever certain functionals on the graph $G$ are themselves bounded.
+One of these functionals will be a measure of eigenvector incoherence.
+	\item We analyze the behavior of these functionals with respect to the random graph $G_{n,r}$, and bound them with high probability.
+	\item We condition on $X \subseteq \mathcal{E}$, where $\mathcal{E}$ is a high-probability set over which the relevant functionals on $G_{n,r}$ are bounded.
+We conclude that the upper bound on testing error derived in our first step holds whenever $X \subseteq \mathcal{E}$.
+\end{enumerate}
+
+\subsubsection{Step 1: Testing error on a fixed graph}
+
+Let $G = (V,E)$ be a graph over vertices $V = \set{v_1,\ldots,v_n}$, and let $\beta^{(p)} = (\beta_1^{(p)},\ldots,\beta_n^{(p)}) \in \Reals^n$, $\beta^{(q)} = (\beta_1^{(q)},\ldots,\beta_n^{(q)}) \in \Reals^n$ be non-negative signals over the vertices $V$.
+We observe labels $a = (a_1,\ldots,a_n)$ according to the model
+\begin{equation*}
+a_i = 
+\begin{cases*}
+1, & \textrm{with probability $\frac{\betap_i}{\betap_i + \betaq_i}$} \\
+-1, & \textrm{with probability $\frac{\betaq_i}{\betap_i + \betaq_i}$}
+\end{cases*}
+\end{equation*}
+Letting $L = VSV^T$ be the spectral decomposition of the Laplacian $L$ of $G$, our graph spectral test statistic is 
+\begin{equation*}
+T_{\spec}^{(2)} = \frac{1}{n}\sum_{k = 1}^{\kappa} \left(\sum_{i = 1}^{n} v_i a_i\right)^2
+\end{equation*}
+where $\kappa$ is a tuning parameter. The resulting test we will use is
+\begin{equation*}
+\phi_{\spec}^{(2)} = \1\{T_{\spec}^{(2)} \geq \frac{\kappa}{n} + t(b)\},~~\textrm{where}~ t(b) = b\sqrt{\frac{2\kappa}{n^2}}~~\textrm{for $b \geq 1$.}
+\end{equation*}
+
+Let $\Pi_{\max}(\kappa;G)$ be a measure of the incoherence of the eigenvectors $v_1,\ldots,v_\kappa$, given by
+\begin{equation*}
+\Pi_{\max}(\kappa;G) := \max_{i = 1,\ldots,n} \left\{\sum_{k = 1}^{\kappa} v_{k,i}^2\right\}
+\end{equation*}
+In Lemma~\ref{lem:twosample_fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}^{(2)}$.
+Our bound on the Type II error will be stated as a function of $\Pi_{\max}(\kappa;G)$ as well as the $\kappa$th eigenvalue $s_{\kappa}$ and the smoothness functional $S_2(\beta;G)$.
+We use the notation $\varDelta^{(p,q)} = (\betap - \betaq)/(\betap + \betaq)$.
+
+\begin{lemma}
+	\label{lem:twosample_fixed_graph_testing}
+	Let $1 \leq \kappa \leq n$ be an integer.
+	\begin{enumerate}
+		\item \textbf{Type I error:} Under the null hypothesis $\betap = \betaq$, the Type I error of $\phi_{\spec}^{(2)}$ is upper bounded
+		\begin{equation}
+		\label{eqn:twosample_graph_spectral_type_I_error}
+		\mathbb{E}_{\betap,\betap}(\phi_{\spec}^{(2)}) \leq \frac{1}{b^2}.
+		\end{equation}
+		\item \textbf{Type II error:} Suppose $\Pi_{\max}(\kappa;G) \leq 1$. Then for any $b$, $\betap$ and $\betaq$ such that
+		\begin{equation}
+		\label{eqn:twosample_fixed_graph_testing_critical_radius}
+		\frac{1}{n} \sum_{i = 1}^{n} (\varDelta_i^{(p,q)})^2 \geq \frac{1}{1 - \Pi_{\max}(\kappa;G)}\left(2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_2(\varDelta;G)}{ns_{\kappa}}\right)
+		\end{equation}
+		the Type II error of $\phi_{\spec}^{(2)}$ is upper bounded,
+		\begin{equation}
+		\label{eqn:twosample_graph_spectral_type_II_error}
+		\mathbb{E}_{\betap,\betaq}(1 - \phi_{\spec}^{(2)}) \leq \frac{3}{b}.
+		\end{equation}
+	\end{enumerate}
+\end{lemma}
+
+\paragraph{Proof of Lemma~\ref{lem:twosample_fixed_graph_testing}.}
+
+To prove Lemma~\ref{lem:twosample_fixed_graph_testing} we will compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}^{(2)}$.
+The inequalities \eqref{eqn:twosample_graph_spectral_type_I_error} and \eqref{eqn:twosample_graph_spectral_type_II_error} can then be computed using Chebyshev's inequality in a manner very similar to the proof of Lemma~\ref{lem:fixed_graph_testing}, and we omit the details.
+Let $w_i := a_i - \vardeltapq_{i}$, and note that $\Ebb(w_i) = 0$ and
+\begin{equation*}
+\Ebb(w_i^2) = \frac{\betap_i \betaq_i}{(\betap_i + \betaq_i)^2} = 1 - (\vardeltapq_{i})^2,~~ \Ebb(w_i^3) = \frac{1}{2}(1 - (\vardeltapq_{i})^2)\vardeltapq_{i}
+\end{equation*}
+
+\vspace{.2 in}
+
+\textit{Mean of $T_{\spec}^{(2)}$}:
+
+We have
+\begin{align*}
+\Ebb(T_{\spec}^{(2)}) & = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{\vardeltapq}{v_k}^2 + \Ebb\bigl(\dotp{w}{v_k}^2 + 2 \dotp{w}{v_k} \dotp{\vardeltapq}{v_k}\bigr)\right) \\
+& = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{1 - (\vardeltapq)^2}{v_k^2} + \dotp{\vardeltapq}{v_k}\right) \\
+& \geq \frac{\kappa}{n} + \left(\sum_{k = 1}^{\kappa}\dotp{\vardeltapq}{v_k} - \dotp{(\vardeltapq)^2}{v_k^2}\right)
+\end{align*}
+When $\vardeltapq = 0$, this equals $\kappa/n$. Otherwise, we have the following upper bound:
+\begin{align*}
+\sum_{k = 1}^{\kappa} \dotp{(\vardeltapq)^2}{v_k^2}  & = \sum_{i = 1}^{n} (\vardeltapq_{i})^2 \left(\sum_{k = 1}^{\kappa} v_{k,i}^2 \right)\\
+& \leq \Pi_{\max}(\kappa,G) \cdot \sum_{i = 1}^{n} (\vardeltapq_{i})^2 = \Pi_{\max}(\kappa,G) \cdot \norm{\vardeltapq}_2^2.
+\end{align*}
+Additionally, $\sum_{k = 1}^{\kappa}\dotp{\vardeltapq}{v_k} \geq \norm{\vardeltapq}_2^2 - S_2(\beta;G)/s_{\kappa}$ by reasoning given in the proof of Lemma~\ref{lem:fixed_graph_testing}, and as a result
+\begin{equation*}
+\Ebb(T_{\spec}^{(2)}) \geq \frac{\kappa}{n} -  \frac{\bigl(1 - \Pi_{\max}(\kappa;G)\bigr)}{n} \norm{\vardeltapq}_2^2 - \frac{S_2(\beta;G)}{ns_{\kappa}}
+\end{equation*}
+
+\vspace{.2 in}
+\textit{Variance of $T_{\spec}^{(2)}$:}
+
+We have
+\begin{align}
+\Var(T_{\spec}^{(2)}) & = \frac{1}{n^2}\Var(2\dotp{V_{\kappa}V_{\kappa}^T\vardeltapq}{w} + \dotp{V_{\kappa}V_{\kappa}^Tw}{w}) \nonumber \\
+& = \frac{1}{n^2}\Bigl\{4\underbrace{\Var(\dotp{V_{\kappa}V_{\kappa}^T\vardeltapq}{w})}_{=: V_1} + 2\underbrace{\Cov(\dotp{V_{\kappa}V_{\kappa}^T\vardeltapq}{w},\dotp{V_{\kappa}V_{\kappa}^Tw}{w})}_{=: K_1} + \underbrace{\Var(\dotp{V_{\kappa}V_{\kappa}^Tw}{w})}_{=:V_2}\Bigr\}, \label{eqn:var}
+\end{align}
+and we now upper bound each of the three terms on the right hand side of the previous display.
+
+\textbf{Upper bound on $V_1$:}
+Let $\Sigma := \Cov(w)$ be the covariance matrix of $w$. Noting that $\Sigma \preceq I$, we have
+\begin{equation}
+\label{eqn:var1}
+\Var(\dotp{V_{\kappa}V_{\kappa}^T\vardeltapq}{w}) =(\vardeltapq)^T V_{\kappa}V_{\kappa}^T \Sigma V_{\kappa}V_{\kappa}^T \vardeltapq \leq \norm{V_{\kappa}V_{\kappa}^T\vardeltapq}^2.
+\end{equation}
+
+\textbf{Upper bound on $K_1$:}
+Noting that $\Ebb(\dotp{V_{\kappa}V_{\kappa}^T \vardeltapq}{w}) = 0$, we have that
+\begin{align}
+K_1 & = \Ebb\left[\dotp{V_{\kappa}V_{\kappa}^T \vardeltapq}{w} \dotp{V_{\kappa}V_{\kappa}^T w}{w}\right] \nonumber \\
+& = \sum_{i = 1}^{n} \sum_{j = 1}^{n} \sum_{i' = 1}^{n} \Ebb\left[w_i w_j (V_{\kappa}V_{\kappa}^T)_{ij} w_{i'} (V_{\kappa}V_{\kappa}^T\vardeltapq)_i\right] \nonumber \\
+& = \sum_{i = 1}^{n} \Ebb\left[w_i^3\right] (V_{\kappa}V_{\kappa}^T)_{ii} (V_{\kappa}V_{\kappa}^T \vardeltapq)_i \nonumber \\
+& = \frac{1}{2}\sum_{i = 1}^{n} (1 - (\vardeltapq_{i})^2)(\vardeltapq_{i}) (V_{\kappa} V_{\kappa}^T)_{ii}  (V_{\kappa}V_{\kappa}^T\vardeltapq)_i \nonumber \\
+& \leq \norm{V_{\kappa}V_{\kappa}^T \vardeltapq} \cdot \left(\sum_{i = 1}^{n} (\vardeltapq_{i})^2 (V_{\kappa} V_{\kappa}^T)_{ii}^2\right)^{1/2} \nonumber \\
+& \leq \Pi(\kappa,G) \cdot \norm{V_{\kappa}V_{\kappa}^T \vardeltapq} \cdot \norm{\vardeltapq}  \label{eqn:var2} 
+\end{align}
+
+\textbf{Upper bound on $V_2$:}
+$V_2$ is a variance of a sum, which we re-express as the sum of covariances:
+\begin{align*}
+V_2 & = \Var(\dotp{V_{\kappa}V_{\kappa}^T w}{w}) \\
+& = \sum_{i,j,i',j' = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ij} (V_{\kappa}V_{\kappa}^T)_{i'j'} \Cov(w_i w_j, w_{i'} w_{j'}).
+\end{align*}
+This covariance will be non-zero only when $i = i' \neq j = j'$, $i = j' \neq j = i'$, or $i = i' = j = j'$, and therefore
+\begin{align}
+V_2 & = 2\sum_{i = 1}^{n}\sum_{j = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ij}^2 \Var(w_iw_j) + \sum_{i = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ii}^2 \Var(w_i^2) \nonumber\\
+& \leq 2\sum_{i = 1}^{n}\sum_{j = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ij}^2 + \sum_{i = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ii}^2 \leq 3 \mathrm{tr}\bigl((V_{\kappa}V_{\kappa}^T)^2\bigr) = 3\kappa.
+\label{eqn:var3}
+\end{align}
+
+Now, plugging \eqref{eqn:var1}, \eqref{eqn:var2}, and \eqref{eqn:var3} back into \eqref{eqn:var}, we obtain
+\begin{equation}
+\label{eqn:var_5}
+\Var(T_{\spec}^{(2)}) \leq \frac{1}{n^2}\left\{4\norm{V_{\kappa}V_{\kappa}^T \vardeltapq}^2 + \Pi(\kappa;G) \cdot \norm{V_{\kappa}V_{\kappa}^T \vardeltapq} \cdot \norm{\vardeltapq} + 3\kappa \right\}.
+\end{equation}
+
+\subsubsection{Step 2: Bounding neighborhood graph functionals}
+
+Note that since $p,q \in W^{1,2}(\mathcal{X};R)$, we have $f := p - q \in W^{1,2}(\mathcal{X};2R)$.
+Therefore the bounds~\eqref{eqn:continuous_to_discrete_sobolev_norm} and \eqref{eqn:l2_to_empirical_norm} apply to $S_2(\varDelta,G_{n,r})$ and $(1/n)\sum_{i = 1}^{n}\varDelta_i^2$, respectively.
+The bound~\eqref{eqn:eigenvalue_tail_bound} continues to apply to $s_{\kappa}$.
+What remains is to upper bound the incoherence parameter
+\begin{equation}
+\label{eqn:pi_max}
+\Pi_{\max}(\kappa,G_{n,r}) \leq \frac{1}{2}~~\textrm{when}~ \kappa = n^{2/5}~\textrm{and}~ r = \log^a n \cdot \left(\frac{\log n}{n}\right)
+\end{equation}
+with probability $1 - o(1)$ as $n \to \infty$.
+
+Our proof of~\eqref{eqn:pi_max} will proceed by relating $G_{n,r}$ to a graph which satsifies a strict notion of incoherence, the chain graph.
+When $G = \overline{G}$, the incoherence parameter $\Pi_{\max}$ is upper bounded
+\begin{equation}
+\Pi_{\max}(\kappa,\overline{G}) \leq \frac{\kappa}{n}.
+\end{equation}
+To prove that the weaker bound \eqref{eqn:pi_max} holds with respect to $G_{n,r}$, we will show that the two graphs $\overline{G}$ and $G_{n,r}$ are $\delta$-spectrally similar, meaning
+\begin{equation}
+\label{eqn:spectral_similarity}
+(1 - \delta_n) x^T L x \leq x^T \overline{L} x \leq (1 + \delta_n) x^T L x~~\textrm{for all}~x \in \Reals^n, \delta_n = \frac{c \log n}{n}
+\end{equation}
+with probability $1 - o(1)$ as $n \to \infty$.
+The proof of~\eqref{eqn:spectral_similarity} relies heavily on bounding the transport distance between $X$ and $\overline{X}$, i.e.
+\eqref{eqn:transport_distance}.
+Then the following technical Lemma along with \eqref{eqn:spectral_similarity} will imply \eqref{eqn:pi_max}.
+\begin{lemma}
+	\label{lem:pi_max_pf_1}
+	Let $G$ satisfy~\eqref{eqn:spectral_similarity} for a given $0 \leq \delta < 1$. Then for any $R \gtrsim n \delta^{1/2} (s_{n}^{1/2} \vee 1)$ and $k \leq n/8$,
+	\begin{equation*}
+	v_{k,i}^2 \lesssim \frac{1}{n}\left(R + \frac{n^4 \delta^2 s_{n}^2}{R^3}\right) \quad \textrm{for every $i = 1,\ldots,n$.}
+	\end{equation*} 
+\end{lemma}
+At a high level, Lemma~\ref{lem:pi_max_pf_1} is proved through repeated applications of the Davis-Kahan Theorem.
+The full proof is long and we delay presenting it until Section~\ref{sec:technical_lemma_proofs}.
+
+\subsubsection{Step 3: Conclusion}
+
+With Lemma~\ref{lem:twosample_fixed_graph_testing} as well as \eqref{eqn:continuous_to_discrete_sobolev_norm}, \eqref{eqn:eigenvalue_tail_bound}, \eqref{eqn:l2_to_empirical_norm} and \eqref{eqn:pi_max} in hand, the proof of Theorem~\ref{thm:twosample_sobolev_testing_rate} follows by similar reasoning to the conclusion of Theorem~\ref{thm:sobolev_testing_rate}.
+
+\section{Proofs of Technical Results}
+\label{sec:technical_lemma_proofs}
+
+Here we give in full detail the proof of \eqref{eqn:spectral_similarity} and Lemma~\ref{lem:pi_max_pf_1}.
+
+\subsection{Proof of~\eqref{eqn:spectral_similarity}}
+
+Assuming~\eqref{eqn:transport_distance} holds and $r$ is chosen according to~\eqref{eqn:pi_max}, we have already shown that the upper bound in~\eqref{eqn:spectral_similarity} holds for $\delta = 0$ (see~\eqref{eqn:eigenvalue_tail_bound_pf1}).
+Now we need only to show the lower bound, which we will do following these steps:
+\begin{enumerate}[(i)]
+	\item First, we consider the case of two fixed graphs $G$ and $\wt{G}$, and establish a Poincare inequality.
+In other words, we show that if edges in $G$ can be mapped to short, non-overlapping paths in $\wt{G}$, then the Laplacian matrix $L$ cannot be much larger $\wt{L}$, in the usual sense of partial ordering of matrices.
+	\item Now, we treat the case of $G = G_{n,r}$ and $\wt{G} = \overline{G}$.
+We assume that~\eqref{eqn:transport_distance} holds, and exhibit such a mapping between edges in $G_{n,r}$ and paths in $\overline{G}$.
+The resulting Poincare inequality will imply~\eqref{eqn:spectral_similarity}.
+\end{enumerate}
+In step (ii) we assume that~\eqref{eqn:transport_distance} holds.
+Since \eqref{eqn:transport_distance} holds with probability at least $1 - c/n$, this will imply that~\eqref{eqn:spectral_similarity} holds with at least the same probability.
+
+\subsubsection{Step 1: Poincare inequality}
+
+Let $G = (V,E)$ and $\wt{G} = (V,\wt{E})$ be two graphs over a common vertex set $V$. The set $\mathcal{P}_{\wt{G}}$ consists of all paths $P$ over $\wt{G}$, meaning all tuples
+\begin{equation*}
+P = (\wt{e}_1,\wt{e}_2,\ldots,\wt{e}_m),~~m \in \mathbb{N}~, \wt{e}_j \in \wt{E}~ \forall~ j \in [m], (\wt{e}_j)_2 = (\wt{e}_{j + 1})_1 \forall~j \in [m - 1].
+\end{equation*}
+Let $\gamma:E \to \mathcal{P}_{\wt{G}}$ be a mapping from edges in $G$ to paths in $\wt{G}$. The maximum path length $M(\gamma)$ is defined
+\begin{equation*}
+M(\gamma) = \max_{e \in E} \abs{\gamma(e)}
+\end{equation*} 
+and the bottleneck $b(\gamma)$ is defined
+\begin{equation*}
+b(\gamma) = \max_{\wt{e} \in \wt{E}} \abs{\{e \in E:  \wt{e} \in \gamma(e)\}}.
+\end{equation*}
+As mentioned previously, the Laplacian matrix $L$ cannot be much larger than $\wt{L}$ if edges in $G$ can be mapped to short, non-overlapping paths in $\wt{G}$.
+Formally, we will show the following: for any $\gamma: E \to \mathcal{P}_{\wt{E}}$,
+\begin{equation}
+\label{eqn:poincare_inequality}
+x^T L x \leq M(\gamma)\cdot b(\gamma) \cdot (x^T \wt{L} x)~\forall{x \in \Reals^n}
+\end{equation}
+
+\textcolor{red}{TODO}: Prove Poincare inequality.
+
+\subsubsection{Step 2: Mapping $E_{G_{n,r}}$ to $\mathcal{P}_{\overline{G}}$}
+We assume there exists a mapping $\pi:\overline{X} \to X$ which satisfies \eqref{eqn:transport_distance}.
+We will apply the Poincare inequality to $\pi^{-1}(G_{n,r})$, the isomorphism of $G_{n,r}$ induced by the mapping $\pi^{-1}: X \to \overline{X}$.
+We now construct our mapping $\gamma:\pi^{-1}(E) \to \overline{E}$.
+For a given $(\overline{x}_{k},\overline{x}_{\ell}) \in \pi^{-1}(E)$, recalling the notation $t = n^{1/d}$, let
+\begin{equation*}
+s_i = \frac{1}{t}(\textrm{sign}(k_i - \ell_i),0,\ldots,0),~~\textrm{for $i = 1,\ldots,d$.}
+\end{equation*}
+and
+\begin{align*}
+\gamma((\overline{x}_{k},\overline{x}_{\ell})) = \bigl(& (\overline{x}_k, \overline{x}_{k} + s_1), (\overline{x}_{k} + s_1,\overline{x}_{k} + 2s_1),\ldots,(\overline{x}_{k} + (\abs{k_1 - \ell_1} - 1)s_1, \overline{x}_{k} + \frac{1}{t}(k_1 - \ell_1)) \\
+& (\overline{x}_{k} + \frac{1}{t}(k_1 - \ell_1), \overline{x}_{k} + \frac{1}{t}(k_1 - \ell_1) + s_2),\ldots,(\overline{x}_{k} + \frac{1}{t}(k_1 - \ell_1) + (\abs{k_2 - \ell_2} - 1)s_2, \overline{x}_{k}  + \frac{1}{t}((k_1 - \ell_1) + (k_2 - \ell_2)) \\
+& \vdots \\
+& (\overline{x}_{k} + \frac{1}{t}\sum_{i = 1}^{d - 1}(k_i - \ell_i) + s_d,\overline{x}_{k} + \frac{1}{t}\sum_{i = 1}^{d - 1}(k_i - \ell_i) + 2s_d), \ldots, (\overline{x}_k + \frac{1}{t}\sum_{i = 1}^{d - 1}(k_i - \ell_i) + (\abs{k_d - \ell_d} - 1)s_d, \overline{x}_{\ell})\bigr)
+\end{align*}
+As a result of~\eqref{eqn:transport_distance} we obtain the following bounds on $M(\gamma)$ and $b(\gamma)$:
+\begin{align}
+M(\gamma) & \leq c n^{1/d} \left(\left(\frac{\log n}{n}\right)^{1/d} + r\right) \label{eqn:maximum_path_length}\\
+b(\gamma) & \leq c \left(n^{1/d} \left(\left(\frac{\log n}{n}\right)^{1/d} + r\right)\right)^{2d}, \label{eqn:bottleneck}
+\end{align}
+which we now prove.
+
+\paragraph{Proof of~\eqref{eqn:maximum_path_length}:}
+
+Suppose $(\overline{x}_k, \overline{x}_{\ell}) \in \pi^{-1}(E)$. Then
+\begin{align*}
+\gamma(\overline{x}_k, \overline{x}_{\ell})) = \norm{\overline{x}_k - \overline{x}_{\ell}}_1 & \leq \sqrt{d} \norm{\overline{x}_k - \overline{x}_{\ell}}_2 \\
+& \leq \sqrt{d}\bigl(\norm{\overline{x}_k - \pi(\ol{x}_{k})}_2 + \norm{\pi(\ol{x}_k) - \pi(\ol{x}_{\ell})}_2 + \norm{\overline{x}_\ell - \pi(\ol{x}_{\ell})}_2\bigr) \\
+& \leq \sqrt{d}\left(\norm{\pi(\ol{x}_k) - \pi(\ol{x}_{\ell})}_2 + 2c\left(\frac{\log n}{n}\right)^{1/d}\right) \tag{by \eqref{eqn:transport_distance}} \\
+& \leq \sqrt{d}\left(r + 2c\left(\frac{\log n}{n}\right)^{1/d}\right) \tag{since $(\pi(\ol{x}_k),\pi(\ol{x}_{\ell})) \in E$}
+\end{align*}
+
+\paragraph{Proof of~\eqref{eqn:bottleneck}:} \textcolor{red}{TODO}:
+
+\subsection{Proof of Lemma~\ref{lem:pi_max_pf_1}}
+
+Write $\overline{L} = \overline{V}\overline{S} \overline{V}^T$ for the spectral decomposition of $\overline{L}$. The proof of Lemma~\ref{lem:pi_max_pf_1} will proceed according to the following steps.
+\begin{enumerate}
+	\item We show that the incoherence $\max_{i} v_{k,i}^2$ can be upper bounded
+	\begin{equation}
+	\label{eqn:incoherence_proof_1}
+	\max_{i} v_{k,i}^2 \leq \frac{1}{n} \left(\sum_{j = 1}^{n} \abs{\dotp{v_k}{\overline{v}_j}}\right)^2
+	\end{equation}
+	replacing the norm $\norm{v_k^2}_{\infty}$ by the inner products $\dotp{v_k}{\overline{v}_j}$.
+	\item Let $I(k,r) = \set{j \geq 0: \abs{j - k} \leq r}$. Using Davis-Kahan, we establish that the following upper bounds
+	\begin{equation}
+	\label{eqn:incoherence_proof_2}
+	\sum_{j \not\in I(k,r)}^{n} (\dotp{v_k}{\overline{v}_j})^2 \leq 400 \frac{n^4 \delta^2 s_{\max}^2}{R^4}
+	\end{equation}
+	hold for each $k \leq n/8$ and any $r \geq 8k\sqrt{\delta}$. 
+	\item We carefully upper bound the $L_1$ norm present in \eqref{eqn:incoherence_proof_1} given the various bounds on $L_2$ norm we've established in \eqref{eqn:incoherence_proof_2}.
+\end{enumerate}
+
+\paragraph{Step 1: Proof of~\eqref{eqn:incoherence_proof_1}.}
+
+We can re-express $v_k$ in the basis $(\overline{v}_j)$ as
+\begin{equation*}
+v_k = \sum_{j = 1}^{n} \dotp{v_k}{\overline{v}_j} \overline{v}_j,
+\end{equation*}
+whereupon the bound~\eqref{eqn:incoherence_proof_1} follows from the incoherence property of the chain
+\begin{equation*}
+\max_{i,k = 1,\ldots,n} \overline{v}_{k,i}^2 \leq \frac{1}{n}.
+\end{equation*}
+\paragraph{Step 2: Proof of~\eqref{eqn:incoherence_proof_2}.}
+
+Let $\widetilde{L} = L + H$. By Davis Kahan, we have that
+\begin{equation}
+\label{eqn:davis_kahan}
+\sum_{j \not\in I(k,r)} (\dotp{v_k}{\overline{v}_j})^2 \leq \left(\frac{\norm{H}_{\textrm{op}}}{\min{\abs{\overline{s}_j - s_k}:j \in I(k,r)}}\right)^2
+\end{equation}
+To upper bound the numerator, we use~\eqref{eqn:spectral_similarity} to get
+\begin{equation*}
+\norm{H}_{op} \leq \delta s_{n}.
+\end{equation*}
+
+To lower bound the denominator, we note
+\begin{align}
+\abs{\overline{s}_j - s_k} & \geq \abs{\overline{s}_j - \overline{s}_k} - \abs{\overline{s}_k - s_k} \nonumber \\
+& \geq \abs{\overline{s}_j - \overline{s}_k} - \delta \overline{s}_k. \label{eqn:incoherence_proof_8}
+\end{align}
+The eigenvalues of the chain graph are well known to be
+\begin{equation*}
+\overline{s}_j = 2\left(1 - \cos\left(\frac{j\pi}{n}\right)\right) = 4\sin^2\left(\frac{k\pi}{n}\right) ~~\textrm{for $k = 0,\ldots,n - 1$.}
+\end{equation*}
+By Taylor expansion, for $k \leq j \leq n/(2\pi)$ we have
+\begin{equation}
+\label{eqn:incoherence_proof_5}
+\overline{s}_j - \overline{s}_k = 2\left(\cos\left(\frac{k\pi}{n}\right) - \cos\left(\frac{j\pi}{n}\right)\right) \geq \frac{(k - j)^2\pi^2}{n^2}.
+\end{equation}
+and when $j \geq n/(2\pi)$ we have $\overline{s}_j - \overline{s}_k \geq 2(\cos\left(\frac{\pi}{8}\right) - \cos\left(\frac{1}{2}\right) > .09$.
+Additionally since $\sin(x) \leq x$ for all $x \geq 0$, we have
+\begin{equation}
+\label{eqn:incoherence_proof_6}
+\overline{s}_k \leq 4\frac{k^2\pi^2}{n^2}.
+\end{equation}
+Combining~\eqref{eqn:incoherence_proof_5},\eqref{eqn:incoherence_proof_6}, and the lower bound $\abs{k - j} \geq R \geq 8k\sqrt{\delta}$, we have that
+\begin{equation}
+\label{eqn:incoherence_proof_7}
+\abs{\overline{s}_j - \overline{s}_k} - \delta \overline{s}_k \geq .04\frac{(k - j)^2\pi^2}{n^2} - 4 \delta\frac{k^2\pi^2}{n^2} \geq .02\frac{(k - j)^2\pi^2}{n^2}.
+\end{equation}
+The result follows from \eqref{eqn:incoherence_proof_7}, \eqref{eqn:incoherence_proof_8} and \eqref{eqn:davis_kahan}.
+
+\paragraph{Step 3: $L_1$ norm to $L_2$ norm.}
+
+Let $u \in \Reals^n$, and suppose we know that the $L_2$ norm of $u$ is bounded,
+\begin{equation}
+\label{eqn:incoherence_proof_3}
+\norm{u}_2^2 \leq 1.
+\end{equation}
+Under no other conditions on $u$, the upper bound $\norm{u}_1 \leq \sqrt{n}$ is the best that can be hoped for (achieved when $u = (n^{-1/2},\ldots,n^{-1/2})$).
+However, suppose we also know that for some $B_2^2 \geq B_3^2 \geq \ldots \geq B_n^2$, we have that there exists some $R \in [n]$ such that
+\begin{equation}
+\label{eqn:incoherence_proof_4}
+\sum_{j = r}^{n} v_j^2 \leq B_r^2 ~~\textrm{for each $r = R,\ldots,n$.}
+\end{equation}
+Clearly, if $B_n^2 \leq \frac{1}{n}$ then $u$ cannot be equal to $(n^{-1/2},\ldots,n^{-1/2})$.
+We might hope for more general improvements on the bound $\norm{u}_1 \leq \sqrt{n}$ if $B_R^2$ is quite small once $R$ gets sufficiently large.
+The following Lemma gives such a result.
+\begin{lemma}
+	\label{lem:pi_max_pf_1_util_1}
+	Suppose $u \in \Reals^n$ satisfies \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4} for some $1 = B_1^2 = B_2^2 \geq \ldots \geq B_n^2 \geq B_{n+1}^2 = 0$.
+Assume additionally that there exists an $R \in [n]$ such that
+	\begin{equation}
+	\label{eqn:incoherence_util_1}
+	B_r^2 - B_{r + 1}^2 \leq \frac{1}{r}~\textrm{for each $r = R, R+1,\ldots, n - 1$.}
+	\end{equation}
+	Then,
+	\begin{equation}
+	\label{eqn:incoherence_util_2}
+	\norm{u}_1 \leq \sqrt{R}\sqrt{1 - B_R^2} + \sum_{j = R}^{n} \sqrt{B_j^2 - B_{j+1}^2}.
+	\end{equation}
+\end{lemma}
+\begin{proof}
+	Set
+	\begin{equation*}
+	(u_\star)_j^2 = 
+	\begin{cases*}
+	\frac{1}{R}(1 - B_R^2), ~ j \leq R \\
+	B_j^2 - B_{j + 1}^2, j > R
+	\end{cases*}
+	\end{equation*}
+	so that $\norm{u_\star}_1$ is equal to the right hand side of~\eqref{eqn:incoherence_util_2}.
+We now prove by contradiction that $u_\star$ maximizes $\norm{\cdot}_1$ under the constraints \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4}.
+Suppose this were not true, and that the maximum is achieved for some $u \neq u_{\star}$.
+	\begin{itemize}
+		\item since $\norm{u}_1 > \norm{u_{\star}}$ then there exists some index $j$ such that
+		\begin{equation}
+		\label{eqn:incoherence_util_3}
+		u_j > (u_{\star})_j
+		\end{equation} 
+		\item by~\eqref{eqn:incoherence_proof_3} $\norm{u}_2 \leq 1$. Since $\norm{u_{\star}}_2 = 1$, by \eqref{eqn:incoherence_util_3} there must exist some index $k$ such that
+		\begin{equation}
+		\label{eqn:incoherence_util_4}
+		(u_{\star})_k > u_k.
+		\end{equation} 
+		Choose $k$ to be the largest of all such indices.
+		\item suppose $k < j$ and $j > R$.
+Since $k$ was chosen to be the largest such index which satisfied~\eqref{eqn:incoherence_util_4}, clearly $v_i \geq (u_{\star})_i$ for all $i > j$, and by \eqref{eqn:incoherence_util_3} $u_j > (u_{\star})_j$.
+But then
+		\begin{equation*}
+		B_j^2 = \sum_{i = j}^{N} (u_{\star})_i^2 < \sum_{i = j}^{N} u_i^2 
+		\end{equation*}
+		and so $v$ does not satisfy~\eqref{eqn:incoherence_proof_4}. Therefore either $k > j$ or $j \leq R$.
+		\item In either case, we have
+		\begin{equation*}
+		u_j > (u_{\star})_j \geq (u_{\star})_k > u_k.
+		\end{equation*}
+		Let $j \leq l < k$ be the largest index for which $v_j > (a_{\star})_j$.
+
+		We now construct a vector $\wt{u}$ which satisfies the constraints \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4} such that $\norm{\wt{u}}_1 > \norm{u}_1$.
+Once we have shown this, we will have established a contradiction, and the proof of Lemma~\ref{lem:pi_max_pf_1_util_1} will be complete.
+Let $\wt{u} = (\wt{u}_i)$ be given as follows:
+		\begin{equation*}
+		\wt{u}_i = 
+		\begin{cases*}
+		u_i, ~~\textrm{if $i \neq k,l$}, \\
+		u_{\star,k}, ~~\textrm{if $i = k$}, \\
+		\sqrt{u_l^2 - (u_{\star,k}^2 - a_k^2)}, ~~\textrm{if $i = l$.}
+		\end{cases*}
+		\end{equation*}
+		Clearly $\norm{\wt{u}}_2 = \norm{u}_2 = 1$ and so $\wt{u}$ satisfies~\eqref{eqn:incoherence_proof_3}.
+To see that $\wt{u}$ satisfies~\eqref{eqn:incoherence_proof_4} we separate the analysis into cases.
+When $r \geq l + 1$, we have that
+		\begin{equation*}
+		\sum_{i = r}^{n} \wt{u}_i^2 = \sum_{i = r}^{n} u_i^2 \leq B_r^2.
+		\end{equation*}
+		When $l < r \leq k$, we have that
+		\begin{equation*}
+		\sum_{i = r}^{n} \wt{u}_i^2 \leq \sum_{i = r}^{k} u_{\star,i}^2 + \sum_{i = k +1}^{n} u_{i}^2 \leq B_r^2 - B_{k + 1}^2 + B_{k + 1}^2 = B_r^2
+		\end{equation*}
+		Finally, when $R \leq r \leq l$, we have
+		\begin{equation*}
+		\sum_{i = r}^{n} \wt{u}_i^2 \leq u_l^2 - (u_{\star,k}^2 - u_k^2) + (u_{\star,k}^2) + \sum_{i \neq k,l} u_i^2 = \sum_{i = r}^{n} u_i^2 \leq B_r^2.
+		\end{equation*}
+		Therefore $\wt{u}$ satisfies~\eqref{eqn:incoherence_proof_4}. Finally, we have that
+		\begin{equation*}
+		\norm{\wt{u}}_1 = \norm{u}_1 + (u_{\star,k}  - u_k) - (\sqrt{u_l^2 - (u_{\star,k}^2 - u_k^2)} - u_l) > \norm{u}_1 + (u_{\star,k}  - u_k) - (u_{\star,k}  - u_k) = \norm{u}_1,
+		\end{equation*}
+		so we have established the desired contradiction.
+	\end{itemize}
+\end{proof}
+
+\paragraph{Putting the pieces together.}
+We apply Lemma~\ref{lem:pi_max_pf_1_util_1} to the vector $u = (u_i)_{i = 1}^{n}$, where
+\begin{equation*}
+u_i = \frac{1}{\sqrt{2}}\left(\abs{\dotp{v_k}{\overline{v}_{k - i}}} + \abs{\dotp{v_k}{\overline{v}_{k + i}}}\right).
+\end{equation*}
+Note the following:
+\begin{itemize}
+	\item $\norm{u}_2 \leq 1$.
+	\item By \eqref{eqn:incoherence_proof_2}, the vector $u$ satisfies the constraint~\eqref{eqn:incoherence_proof_4} with
+	\begin{equation*}
+	B_r^2 = 400\frac{n^4\delta^2s_{n}^2}{r^4} ~~\textrm{when $r \geq 8 k \sqrt{\delta}$.}
+	\end{equation*}
+	\item Taylor expanding $f(r + 1) = (r + 1)^{-4}$ around $r$, we have that
+	\begin{equation*}
+	B_r^2 - B_{r + 1}^2 \leq 3200 \frac{n^4 \delta^2 s_{n}^2}{r^5} ~~\textrm{when $r \geq 2$.}
+	\end{equation*}
+	Therefore, the sequence $(B_r)$ satisfies \eqref{eqn:incoherence_util_1} for all $r$ large enough such that 
+	\begin{equation*}
+	r \geq 2^{5/4} 100^{1/4} n \sqrt{\delta s_{n}}
+	\end{equation*}
+	\item Since $u$ and $(B_r)$ satisfy the constraints \eqref{eqn:incoherence_proof_3}, \eqref{eqn:incoherence_proof_4} and \eqref{eqn:incoherence_proof_5}, we may apply Lemma~\ref{lem:pi_max_pf_1_util_1} to $u$, obtaining that
+	\begin{equation*}
+	\sum_{j = 1}^{n} \abs{\dotp{v_j}{\overline{v}_k}} = \sqrt{2}\norm{u}_1 \leq \sqrt{2R} + 60\sqrt{2} n^2 \delta s_{n} \sum_{r = R}^{n} r^{-5/2}
+	\end{equation*}
+	for any $R \geq 2 \vee k\sqrt{\delta} \vee 2^{5/4} 100^{1/4} n \sqrt{\delta s_{n}}$. Bounding sum by integral, we have that if $R \geq 2$ then
+	\begin{equation*}
+	\sum_{r = R}^{n} r^{-5/2} \leq \int_{R - 1}^{n - 1} x^{-5/2} \,dx \leq \frac{2^{3/2}2}{3} R^{-3/2}.
+	\end{equation*}
+	Plugging this into the previous expression, we arrive at
+	\begin{equation*}
+	\sum_{j = 1}^{n} \abs{\dotp{\overline{v}_j}{v_k}} \leq \sqrt{R} + 60 \sqrt{2} n^2 \delta s_{\max} \frac{2^{3/2}2}{3R^{3/2}}
+	\end{equation*}
+	for any $R \geq 2 \vee k\sqrt{\delta} \vee 2^{5/4} 100^{1/4} n \sqrt{\delta s_{\max}}$. Lemma~\ref{lem:pi_max_pf_1} then follows from~\eqref{eqn:incoherence_proof_1}.
+\end{itemize}
+
 \section{Auxiliary Results}
@@
 	\label{lem:paley_zygmund}
-	Let $0 < p < q$, and let $Z$ be a non-negative random variable such that $\mathbb{E}(Z^q) < \infty$. Then, for all $0 \leq \lambda \leq 1$,
+	Let $0 < p < q$, and let $Z$ be a non-negative random variable such that $\mathbb{E}(Z^q) < \infty$. Then, for all $0 \leq \overline{s} \leq 1$,
 	\begin{equation}
 	\label{eqn:paley_zygmund}
-	\mathbb{P}(Z > \lambda \mathbb{E}(Z^p)) \geq \left[(1 - \lambda^p) \frac{\mathbb{E}(Z^p)}{(\mathbb{E}(Z^q))^{p/q}}\right]^{\frac{q}{q - p}}
+	\mathbb{P}(Z > \overline{s} \mathbb{E}(Z^p)) \geq \left[(1 - \overline{s}^p) \frac{\mathbb{E}(Z^p)}{(\mathbb{E}(Z^q))^{p/q}}\right]^{\frac{q}{q - p}}
 	\end{equation}

### 2019-12-20 notes/graph_testing.tex
added notes for the week

@@
 
-Throughout, let $X = \{x_1,
+We establish some notation that we will use throughout. Let $X = \{x_1,
 \ldots, x_n\}$ be a sample drawn i.i.d. from a distribution $P$ on $\Rd$,
@@
 \end{equation*}
-The worst-case risk may be quite close to $1$ unless we require some separation between null and alternative spaces. A more realistic measure of performance is therefore
+The worst-case risk may be quite close to $1$ unless we enforce some separation between null and alternative spaces. A more realistic measure of performance is therefore
 \begin{equation*}
@@
 \label{eqn:graph_spectral_projections}
-T_{\mathrm{spec}} := \frac{1}{n} \sum_{k = 0}^{\kappa} \left(\sum_{i = 1} v_i y_i\right)^2
+T_{\mathrm{spec}} := \frac{1}{n} \sum_{k = 0}^{\kappa} \left(\sum_{i = 1}^{n} v_{k,i} y_i\right)^2
 \end{equation}
@@
 
-\textcolor{red}{WARNING:} While I have proved all the parts I believe are required for the below theorem, I haven't actually put together the pieces yet.
-
 In Theorem~\ref{thm:twosample_sobolev_testing_rate} we show that under some typical regularity conditions on $P$, the test $\phi_{\textrm{spec}}^{(2)} := \1\{T_{\mathrm{spec}}^{(2)} \geq \tau\}$ is, up to log factors, a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};R)$ when $d = 1$.
@@
 
-For a given $s > 0$, the $s$th Holder norm is given by
-\begin{equation*}
-\norm{f}_{C_d^{s}(\mathcal{X})} := \sum_{\abs{\alpha} \leq s} \norm{D^{\alpha}f}_{\infty} + \sum_{\abs{\alpha} = s} \sup_{x,y \in \mathcal{X}} \frac{\abs{D^{\alpha}f(y) - D^{\alpha}f(x)}}{\norm{x - y}_2}
-\end{equation*}
-and the $s$th Holder space $C_d^{s}(\mathcal{X})$ consists of all functions which are $s$ times continuously differentiable with finite $s$ Holder norm.
+For a given $s > 0$, let $\ell = \floor{s}$ be the largest integer strictly less than $s$. Then the $s$th Holder norm is given by
+\begin{equation*}
+\norm{f}_{C_d^{s}(\mathcal{X})} := \sum_{\abs{\alpha} < s} \norm{D^{\alpha}f}_{\infty} + \sum_{\abs{\alpha} = \ell} \sup_{x,y \in \mathcal{X}} \frac{\abs{D^{\alpha}f(y) - D^{\alpha}f(x)}}{\norm{x - y}_2^{s - \ell}}
+\end{equation*}
+and the $s$th Holder space $C_d^{s}(\mathcal{X})$ consists of all functions which are $\ell$ times continuously differentiable with finite Holder norm $\norm{\cdot}_{C_d^{s}(\mathcal{X})}$.
 Denote the Holder ball by $C_d^{s}(\mathcal{X},R) = \set{f \in C_d^{s}(\mathcal{X}): \norm{f}_{C_d^{s}(\mathcal{X})} \leq R}$.
@@
 \end{equation*}
+
+\section{\textcolor{red}{In Progress}}
+
+We say $K_{r}$ is an order-$s$ kernel when 
+\begin{equation*}
+\int x^j K_{r}(x) \,dx = 0, \textrm{for $j = 1,\dots,s$}
+\end{equation*}
+
+Our goal is to prove the following theorem.
+
+\begin{theorem}
+	\label{thm:higher_order_sobolev_testing_rate}
+	Let $b,s,d \geq 1$ and $a > 0$ be fixed constants such that $d < 4s$.
+Suppose that $\Pbb$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p$ bounded above and below by constants, i.e
+	\begin{equation*}
+	0 < p_{\min} < p(x) < p_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
+	\end{equation*}
+	Then the following statement holds: if the test $\phi_{\spec}$ is performed with parameter choices 
+	\begin{equation*}
+	\textrm{$K_r$ an order-$s$ kernel with}~ r = \log^a n \cdot \left(\frac{\log n}{n}\right)^{1/d}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	\end{equation*}
+	then there exists a constant $c$ which may depend on $d,R,p_{\max},b$ and $a$ but is independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
+	\begin{equation}
+	\label{eqn:higher_order_sobolev_testing_rate}
+	\epsilon^2 \geq c \cdot b \cdot n^{-4/(4s + d)} (\log n)^{h(a,d)}
+	\end{equation}
+	the worst-case risk is upper bounded
+	\begin{equation}
+	\label{eqn:higher_order_sobolev_testing_rate_1}
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; \mathcal{W}^{s,2}(\mathcal{X};R)) \lesssim \frac{1}{b}.
+	\end{equation}
+\end{theorem}
+
+\section{\textcolor{red}{In Progress Proofs}}
+
+\subsection{Proof of Theorem~\ref{thm:higher_order_sobolev_testing_rate}}
+
+Let $G = (V,E)$ be a graph over vertices $V = \set{v_1,\ldots,v_n}$, and let $\beta = (\beta_1,\ldots,\beta_n) \in \Reals^n$ be a signal over the vertices $V$.
+We observe responses $Y = (y_1,\ldots,y_n)$ according to the model
+\begin{equation*}
+y_i = \beta_i + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
+\end{equation*}
+Letting $L = VSV^T$ be the spectral decomposition of the Laplacian $L$ of $G$, our graph spectral test statistic is 
+\begin{equation*}
+T_{\spec} = \frac{1}{n}\sum_{k = 1}^{\kappa} \left(\sum_{i = 1}^{n} v_i y_i\right)^2
+\end{equation*}
+where $\kappa$ is a tuning parameter. The resulting test we will use is
+\begin{equation*}
+\phi_{\spec} = \1\{T_{\spec} \geq \frac{\kappa}{n} + t(b)\},~~\textrm{where}~ t(b) = b\sqrt{\frac{2\kappa}{n^2}}~~\textrm{for $b \geq 1$.}
+\end{equation*}
+
+Let $S_s(\beta;G)$ be a measure of smoothness the signal $\beta$ displays over the graph $G$, given by
+\begin{equation*}
+S_s(\beta;G) := \beta^T L^s \beta
+\end{equation*}
+In Lemma~\ref{lem:higher_order_fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}$.
+Our bound on the Type II error will be stated as a function of $S_2(\beta;G)$ as well as the $\kappa$th eigenvalue $s_{\kappa}$.
+
+\begin{lemma}
+	\label{lem:higher_order_fixed_graph_testing}
+	Let $1 \leq \kappa \leq n$ be an integer.
+	\begin{enumerate}
+		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}$ is upper bounded
+		\begin{equation}
+		\label{eqn:higher_order_graph_spectral_type_I_error}
+		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{1}{b^2}.
+		\end{equation}
+		\item \textbf{Type II error:} For any $b$ and $\beta$ such that
+		\begin{equation}
+		\label{eqn:higher_order_fixed_graph_testing_critical_radius}
+		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_s(\beta;G)}{ns_{\kappa}^s}
+		\end{equation}
+		the Type II error of $\phi_{\spec}$ is upper bounded,
+		\begin{equation}
+		\label{eqn:higher_order_graph_spectral_type_II_error}
+		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{3}{b}.
+		\end{equation}
+	\end{enumerate}
+\end{lemma}
+
+To prove Lemma~\ref{lem:higher_order_fixed_graph_testing} we will first compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}$, and then use Chebyshev's inequality to show \eqref{eqn:higher_order_graph_spectral_type_I_error} and \eqref{eqn:higher_order_graph_spectral_type_II_error}.
+
+\vspace{.2 in}
+
+\textit{Mean of $T_{\mathrm{spec}}$:} Using the notation $\dotp{v}{w} = \sum_{i = 1}^{n} v_iw_i$, we have
+\begin{align*}
+\Ebb(T_{\spec}) & = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{\beta}{v_k}^2 + \Ebb\bigl( \dotp{\varepsilon}{v_k}^2 + 2 \dotp{\varepsilon}{v_k} \dotp{\beta}{v_k}\bigr)\right) \\
+& = \frac{\kappa}{n} + \frac{1}{n}\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2.
+\end{align*}
+When $\beta = 0$, this equals $\kappa/n$. Otherwise, we have the following lower bound:
+\begin{align*}
+\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2 & = \norm{\beta}_2^2 - \sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \\
+& \geq \norm{\beta}_2^2 - \frac{1}{s_{\kappa}^s}\sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 s_k^s \\
+& \geq \norm{\beta}_2^2 - \frac{S_s(\beta;G)}{s_{\kappa}^s},
+\end{align*}
+and therefore $\Ebb(T_{\spec}) \geq \kappa/n + n^{-1}(\norm{\beta}_2^2 - S_s(\beta;G)/s_{\kappa}^s)$.
+
+\vspace{.2 in}
+
+\textit{Variance of $T_{\mathrm{spec}}$:}
+We write $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvectors $v_1,\ldots,v_{\kappa}$ as columns. Consequently,
+\begin{align}
+\Var(T_{\spec}) & = \frac{1}{n^2} \Var(y^T V_{\kappa} V_{\kappa}^T y) \\
+& = \frac{1}{n^2} \Var((\beta + \varepsilon)^T V_{\kappa} V_{\kappa}^T (\beta + \varepsilon)) \\
+& = \frac{1}{n^2} \Var(2 \beta^T V_{\kappa} V_{\kappa}^T \varepsilon + \varepsilon^T V_{\kappa} V_{\kappa}^T \varepsilon) \\
+& \leq \frac{1}{n^2}(4 \beta^T V_{\kappa} V_{\kappa}^T \beta + 2\kappa)
+\end{align}
+where the last inequality follows from standard properties of the Gaussian distribution.
+We now move on to showing the desired inequalities \eqref{eqn:higher_order_graph_spectral_type_I_error} and \eqref{eqn:higher_order_graph_spectral_type_II_error}.
+
+\vspace{.2 in}
+
+\textit{Proof of~\eqref{eqn:higher_order_graph_spectral_type_I_error}:} By Chebyshev's inequality,
+\begin{align*}
+\Pbb_{\beta = 0}\bigl(T_{\spec} \geq \frac{\kappa}{n} + t(b)\bigr)
+& \leq \Pbb_{\beta = 0}\bigl(\abs{T_{\spec} - \frac{\kappa}{n}} \geq t(b)\bigr) \\
+& \leq \frac{\Var_{\beta = 0}(T_{\spec})}{t(b)^2} = \frac{1}{b^2}.
+\end{align*}
+
+\vspace{.2 in}
+
+\textit{Proof of~\eqref{eqn:higher_order_graph_spectral_type_II_error}:} For simplicity, we introduce the notation
+\begin{equation*}
+\Delta = \frac{\norm{\beta}_2^2}{n} - \frac{S_s(\beta;G)}{ns_{\kappa}^s}.
+\end{equation*}
+Assumption~\eqref{eqn:higher_order_fixed_graph_testing_critical_radius} implies $\Delta \geq 2 t(b)$. Then another application of Chebyshev's inequality gives us
+\begin{align*}
+\Pbb_{\beta}\bigl(T_{\spec} \leq \frac{\kappa}{n} + t(b)\bigr) & = \Pbb_{\beta}\bigl(T_{\spec} - \Ebb_{\beta}(T_{\spec}) \leq t(b) - \Delta \bigr) \\
+& \leq \Pbb_{\beta}\bigl(\abs{T_{\spec} - \Ebb_{\beta}(T_{\spec})} \leq \Delta - t(b) \bigr) \tag{since $\Delta \geq t(b)$}	\\
+& \leq \frac{\Var_{\beta}(T_{\spec})}{(\Delta - t(b))^2} \\
+& \leq 4\frac{\Var_{\beta}(T_{\spec})}{\Delta^2} \tag{since $\Delta \geq 2t(b)$} \\
+& \leq 4\frac{2\kappa/n^2 + 4\beta^T V_{\kappa} V_{\kappa}^T \beta /n^2}{\Delta^2}.
+\end{align*}
+We handle each summand in the numerator separately. For the first term, since $\Delta \geq 2 t(b)$, we have
+\begin{equation}
+\label{eqn:higher_order_spectral_type_II_error_pf1}
+\frac{2\kappa}{n^2\Delta^2} \leq \frac{1}{2b^2}.
+\end{equation}
+
+For the second term, noting that $\Delta = \beta^T V_{\kappa} V_{\kappa}^T \beta/n$, we have
+\begin{align}
+\frac{\beta^T V_{\kappa} V_{\kappa}^T \beta/n^2}{\Delta^2} & = \frac{1}{n\Delta} \nonumber \\
+& \leq \frac{1}{2nt(b)} \nonumber \\
+& = \frac{1}{2b\sqrt{2\kappa}}, \label{eqn:higher_order_spectral_type_II_error_pf2}
+\end{align}
+and combining~\eqref{eqn:higher_order_spectral_type_II_error_pf1} and~\eqref{eqn:higher_order_spectral_type_II_error_pf2} yields~\eqref{eqn:higher_order_graph_spectral_type_II_error}.
+
+\subsubsection{Step 2: Bounding neighborhood graph functionals}
+
+To make use of Lemma~\ref{lem:fixed_graph_testing} we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$.
+In particular, we will show that for some constants $c_1,c_2,c_3$, the following statements:
+\begin{enumerate}[label=(E\arabic*)]
+	\item 
+	\label{event:higher_order_discrete_sobolev_norm}
+	\textbf{Graph Sobolev norm:} When $\beta = (f(x_1),\ldots,f(x_n))$ for $f \in W^{s,2}(\mathcal{X};R)$, and $r = n^{-1/(2 + d)}$,
+	\begin{equation}
+	\label{eqn:higher_order_continuous_to_discrete_sobolev_norm}
+	S_s(\beta;G_{n,r}) \leq c_1 \cdot n^{2 + s} r^{s(d + 2)} 
+	\end{equation}
+	\item 
+	\label{event:higher_order_eigenvalue_tail_decay}
+	\textbf{Eigenvalue tail bound:} For any $r \geq n^{-1/(2 + d)}$ and for all $\kappa \in [n]$:
+	\begin{equation}
+	\label{eqn:higher_ordereigenvalue_tail_bound}
+	s_{\kappa} \geq c_2 \cdot n r^{d + 2} \kappa^{2/d}
+	\end{equation}
+	\item 
+	\label{event:higher_order_l2_norm}
+	\textbf{Empirical norm of $f$:} When $f \in W^{s,2}(\mathcal{X};R)$ and \textcolor{red}{some other condition}
+	\begin{equation}
+	\label{eqn:higher_order_l2_to_empirical_norm}
+	\frac{1}{n}\sum_{i = 1}^{n}\beta_i^2 = \norm{f}_n^2 \geq c_3 \cdot \norm{f}_{\Leb^2}^2
+	\end{equation}
+\end{enumerate}
+
+each hold with probability $1 - o(1)$ as $n \to \infty$.
 
@@
 We now move on to showing the desired inequalities \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.
-For simplicity, we introduce the notation
-\begin{equation*}
-\Delta = \norm{\beta}_2^2 - \frac{S_2(\beta;G)}{s_{\kappa}}.
-\end{equation*}
 
@@
 
-\textit{Proof of~\eqref{eqn:graph_spectral_type_II_error}:} Assumption~\eqref{eqn:fixed_graph_testing_critical_radius} implies
-\begin{equation*}
-\Delta \geq \frac{\norm{\beta}_2^2}{n} - \frac{C_n^2}{nC^2} \geq 2 t(b).
-\end{equation*}
-Then another application of Chebyshev's inequality gives us
+\textit{Proof of~\eqref{eqn:graph_spectral_type_II_error}:} For simplicity, we introduce the notation
+\begin{equation*}
+\Delta = \frac{\norm{\beta}_2^2}{n} - \frac{S_2(\beta;G)}{ns_{\kappa}}.
+\end{equation*}
+Assumption~\eqref{eqn:fixed_graph_testing_critical_radius} implies $\Delta \geq 2 t(b)$. Then another application of Chebyshev's inequality gives us
 \begin{align*}
@@
 & \leq 4\frac{\Var_{\beta}(T_{\spec})}{\Delta^2} \tag{since $\Delta \geq 2t(b)$} \\
-& \leq 4\frac{2\kappa/n^2 + 4\beta^T V_{\kappa} V_{\kappa}^T \beta /n}{\Delta^2}.
+& \leq 4\frac{2\kappa/n^2 + 4\beta^T V_{\kappa} V_{\kappa}^T \beta /n^2}{\Delta^2}.
 \end{align*}
@@
 
-For the second term, noting that $\Delta = \beta^T V_{\kappa} V_{\kappa}^T \beta$, we have
+For the second term, noting that $\Delta = \beta^T V_{\kappa} V_{\kappa}^T \beta/n$, we have
 \begin{align}
-\frac{\beta^T V_{\kappa} V_{\kappa}^T \beta/n}{\Delta^2} & \leq \frac{1}{n\Delta} \nonumber \\
+\frac{\beta^T V_{\kappa} V_{\kappa}^T \beta/n^2}{\Delta^2} & = \frac{1}{n\Delta} \nonumber \\
 & \leq \frac{1}{2nt(b)} \nonumber \\
@@
 \label{eqn:spectral_similarity}
-(1 - \delta_n) x^T L x \leq x^T \overline{L} x \leq (1 + \delta_n) x^T L x~~\textrm{for all}~x \in \Reals^n, \delta_n = \frac{c \log n}{n}
+(1 - \delta_n) x^T L x \leq x^T \overline{L} x \leq (1 + \delta_n) x^T L x~~\textrm{for all}~x \in \Reals^n, \delta_n = c \cdot (\log n)^{h(a,1) + da}\cdot \frac{\log n}{n}
 \end{equation}
@@
 \end{lemma}
+
+\textcolor{red}{TODO}: Explain how Lemma~\ref{lem:pi_max_pf_1} implies~\eqref{eqn:pi_max}.
+
 At a high level, Lemma~\ref{lem:pi_max_pf_1} is proved through repeated applications of the Davis-Kahan Theorem.
@@
 
-\textcolor{red}{TODO}: Prove Poincare inequality.
+\paragraph{Proof of~\eqref{eqn:poincare_inequality}:}
+
+Letting $c \in \Reals$, we will use the notation $G \preceq c \cdot  \wt{G}$ as shorthand for
+\begin{equation*}
+x^T L x \leq c \cdot  x^T \wt{L} x,~~\textrm{for all $x \in \Reals^n$}.
+\end{equation*}
+
+Let $G_e = (V, \set{e})$ and $P_e = (V, \set{\widetilde{e}: \widetilde{e} \in \gamma(e)})$ be the graphs associated with $e$ and $\gamma(e)$, respectively. By Lemma \ref{lem: path_poincare}, we have
+\begin{equation*}
+G_{e} \preceq \abs{P_e} P_e
+\end{equation*}
+Summing over all $e \in E_G$, we obtain
+\begin{align*}
+G & \preceq \sum_{e \in E_G} \abs{P_e} P_e \\
+& \preceq \max_{e \in E_G} \abs{\gamma(e)} \sum_{e \in E_G} P_e \\
+& \preceq \max_{e \in E_G} \abs{\gamma(e)} b_{\gamma}\cdot \widetilde{G}
+\end{align*}
 
@@
 We assume there exists a mapping $\pi:\overline{X} \to X$ which satisfies \eqref{eqn:transport_distance}.
-We will apply the Poincare inequality to $\pi^{-1}(G_{n,r})$, the isomorphism of $G_{n,r}$ induced by the mapping $\pi^{-1}: X \to \overline{X}$.
-We now construct our mapping $\gamma:\pi^{-1}(E) \to \overline{E}$.
-For a given $(\overline{x}_{k},\overline{x}_{\ell}) \in \pi^{-1}(E)$, recalling the notation $t = n^{1/d}$, let
-\begin{equation*}
-s_i = \frac{1}{t}(\textrm{sign}(k_i - \ell_i),0,\ldots,0),~~\textrm{for $i = 1,\ldots,d$.}
-\end{equation*}
-and
+We will apply the Poincare inequality to $\pi^{-1}(G_{n,r})$, the isomorphism of $G_{n,r}$ induced by the mapping of vertices $\pi^{-1}: X \to \overline{X}$.
+We will map a given $(\overline{x}_{k},\overline{x}_{\ell}) \in \pi^{-1}(E)$ to a shortest path $P \in \mathcal{P}_{\overline{G}}$ (measured by Manhattan distance) between $\overline{x}_k$ and $\overline{x}_{\ell}$.
+Recalling the notation $t = n^{1/d}$ and letting
+\begin{equation*}
+s_i = \frac{1}{t}(\textrm{sign}(k_i - \ell_i),0,\ldots,0),~~\textrm{for $i = 1,\ldots,d$,}
+\end{equation*}
+our mapping $\gamma: \pi^{-1}(E) \to \mathcal{P}_{\overline{G}}$ is given by
 \begin{align*}
@@
 
-\paragraph{Proof of~\eqref{eqn:bottleneck}:} \textcolor{red}{TODO}:
+\paragraph{Proof of~\eqref{eqn:bottleneck}:}
+
+Suppose $(\overline{x}_k, \overline{x}_{\ell}) \in \gamma(e)$ for some $e = (\overline{x}_i, \overline{x}_j) \in \pi^{-1}(E)$.  Then, by the construction of $\gamma$,
+\begin{equation}
+\label{eqn:bottleneck_pf_1}
+\norm{\overline{x}_k - \overline{x}_i}_2 \wedge \norm{\overline{x}_\ell - \overline{x}_j}_2  \leq \norm{\overline{x}_j - \overline{x}_i}_2
+\end{equation}
+As shown in the preceding paragraph, assuming~\eqref{eqn:transport_distance} we have
+\begin{equation*}
+\norm{\overline{x}_j - \overline{x}_i}_2 \leq r + 2c\left(\frac{\log n}{n}\right)^{1/d}
+\end{equation*}
+and therefore by~\eqref{eqn:bottleneck_pf_1},
+\begin{align*}
+\abs{\set{e \in E: (\overline{x}_k, \overline{x}_{\ell}) \in \gamma(e)}} & \leq \abs{\set{\overline{x}_i: \norm{\overline{x}_k - \overline{x}_i}_2 \leq r + 2c\left(\frac{\log n}{n}\right)^{1/d}}} \cdot \abs{\set{\overline{x}_i: \norm{\overline{x}_k - \overline{x}_i}_2 \leq r + 2c\left(\frac{\log n}{n}\right)^{1/d}}} \\
+& \leq \left(n^{1/d}\left( r + 2c\left(\frac{\log n}{n}\right)^{1/d}\right)\right)^2.
+\end{align*}
+
+Putting the pieces together, we have that when $d = 1$ and $r = (\log n)^a (\log n/n)$, then
+\begin{equation*}
+G_{n,r} \preceq c \cdot (\log n)^{(1/d + a) + 2 + 2da}\cdot \overline{G}
+\end{equation*}
+with probability at least $1 - o(1)$.
 
@@
 
+\begin{lemma}[Poincare inequality for path graphs.]
+	\label{lem: path_poincare}
+	Fix $m \geq 0$. For vertices $V = \set{1, \ldots,m}$ define the path $P(1 \to m) = ((1,2),(2,3),\ldots, (m-1,m))$ and $G_{(1,m)}$ to be the graph consisting only of an edge between $1$ and $m$. Then,
+	\begin{equation*}
+	(m - 1) \cdot P(1 \to m) \succeq G_{(1,m)}
+	\end{equation*}
+\end{lemma}
+
 \end{document}

### 2019-12-20 notes/weekly_notes/12.20.19.NotesForWeek.tex
added notes for the week

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{\mathcal{L}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\Unq}{\mathrm{Unq}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 12/7/19 - 12/20/19}
\author{Alden Green}
\date{\today}
\maketitle

We observe random design and responses $(x_i,y_i)$ according to the following regression model: $x_1,\ldots,x_n$ are drawn i.i.d from distribution $P$ with density $p$ supported on $\mathcal{X} \subset \Reals^d$, and
\begin{equation*}
y_i = f(x_i) + \varepsilon_i,~ \varepsilon_i \overset{i.i.d}{\sim} \mathcal{N}(0,1).
\end{equation*}
Suppose we form the undirected, weighted graph $G_{n,r} = (X,W)$ with edge weights $W = (W_{ij})$ formed according to the kernel function $K$ as follows:
\begin{equation*}
W_{ij} = K_r(x_i,x_j) := \frac{1}{r^d}K(\norm{x_i - x_j}^2)
\end{equation*}
Let $L_n$ be the graph Laplacian operator associated with $G_{n,r}$, defined by the action
\begin{equation*}
L_nf(x) := \frac{1}{nr^2} \sum_{i = 1}^{n}(f(x_i) - f(x))K_r(x_i,x).
\end{equation*}
The graph Laplacian $L_n$ induces a class of roughness functionals $R_s(f)$, defined by
\begin{equation*}
R_{s,n}(f) = R_s(f;G_{n,r}) = \frac{1}{n} f^T L_n^s f
\end{equation*} 
We refer to $R_{s,n}(f)$ as the order-$s$ roughness functional. As part of our graph testing work, we have shown that if $f \neq 0$, then for any design points $X$ and resulting neighborhood graph $G$ with Laplacian $L$ such that
\begin{equation}
\label{eqn:bias_variance}
\norm{f}_n^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{R_{s}(f;G) }{\lambda_{\kappa}^s}
\end{equation}
the graph spectral test $\phi_{\spec}$ makes a Type II error with probability at most $3/b$. Suppose
\begin{equation}
\label{eqn:graph_functionals}
\norm{f}_n^2 \geq \frac{1}{2}\norm{f}_2^2, ~R_s(f;G_{n,r}) \leq ||f||_{W_{d}^{s}(\Xset)}^2 ~~\textrm{and}~~ \lambda_{\kappa} \leq \kappa^{2/d};
\end{equation}
then it can be verified that choosing $\kappa = n^{-2d/(4s + d)}$, the equation~\eqref{eqn:bias_variance} is satisfied whenever $\norm{f}_2^2 \geq n^{-4s/(4s + d)}$. It is therefore sufficient to show that~\eqref{eqn:graph_functionals} holds with high probability over the random design points $X$.

In this week's notes, we will focus our attention on upper bounding the roughness functional $R_{n,s}(f)$. Our main results should take the following form: when (a) $f \in C^{s}(L)$ for a fixed constant $L$ (or more ideally $f \in \mathcal{W}^{s}(L)$), (b) the neighborhood graph radius $r = r(n)$ is properly chosen as a function of $n$, and (c) $K$ is an appropriately chosen kernel, the roughness functional satisfies
\begin{equation}
\label{eqn:roughness_functional_bound}
\mathbb{E}\left[R_{s,n}(f)\right] = O\left(L^2\right)
\end{equation}

These notes detail our current progress in showing such a result for different values of $s$. We showed the following useful representation of $R_{s,n}(f)$ in the \emph{8.7.19} Notes. For $k = (k_1,\ldots,k_n) \in [n]^s$, recursively defined the graph difference operator $D_k$ by 
\begin{equation*}
D_{k_1}f(x) = (f(x_{k_1}) - f(x))K_r(x_{k_1},x),~~ D_kf(x) = \bigl(D_{k_1}(D_{(k_2,\ldots,k_n)}f\bigr)(x)
\end{equation*}
Then when $s$ is even, letting $q = s/2$ we have
\begin{equation}
\label{eqn:roughness_functional_representation_even}
R_{s,n}(f) = \frac{1}{n}\sum_{i = 1}^{n} \left(\frac{1}{(nr^2)^q}\sum_{k \in [n]^q} D_kf(x_i)\right)^2
\end{equation}
and when $s$ is odd, letting $q = (s - 1)/2$ we have
\begin{equation}
\label{eqn:roughness_functional_representation_odd}
R_{s,n}(f) =  \frac{1}{2n^2r^2}\sum_{i,j = 1}^{n}\left(\frac{1}{(nr^2)^q}\sum_{k \in [n]^q}\bigl(D_kf(x_i) - D_kf(x_j)\bigr)\right)^2K_r(x_i,x_j)
\end{equation}
We have already obtained the desired bound on the 1st-order roughness functional $R_{1,n}(f)$, for any $f \in W_d^{1}(L)$, all scalings $r = r(n)$, and the uniform kernel. We therefore begin these notes with the case $s = 2$.

\section{Bounding the 2nd-Order Roughness Functional}
The (first and) second order roughness functionals are somewhat special. These roughness functionals do not involve taking compositions of the difference operator $D_k$, and as a result, we get the desired bounds in expectation even when $r(n)$ is small. 
\begin{lemma}
	\label{lem:2nd_order_roughness_functional}
	Fix $L > 0$, and form the neighborhood graph $G_{n,r}$ using kernel $K(z) = \1\{z^2 \leq r^2\}$. Suppose $f \in C^2(L)$ and $p \in C^1(L)$. Then for any $r(n) \geq n^{-1/(2+d)}$ and , we have
	\begin{equation*}
	\Ebb(R_{2,n}(f)) = O(L^2)
	\end{equation*}
\end{lemma}

\section{Bounding the 3rd-Order Roughness Functional}

For the order-$s$ roughness functionals when $s \geq 3$, the graph Laplacian is forced to approximate higher order derivatives by composing difference operators-- that is, taking differences of differences, etc. In for the roughness functional to continue scaling at the desired rate, we will need $r(n)$ to be larger, and we cannot use the uniform kernel $K$ any longer. In the particular case $s = 3$, we will require a kernel $K$ which is bounded and compactly supported on $B(0,1)$, and additionally satisfies
\begin{equation*}
\int z^j K(z) = 0,~~\textrm{$j = 1,2$.}
\end{equation*}
We then have the following result.
\begin{lemma}
	\label{lem:3rd_order_roughness_functional}
	Suppose $f \in C^{3}(L)$ for some $L > 0$, and further suppose $p \in C^{2}(L)$. For any kernel $K$ which satisfies the previous conditions, and any $r(n) \geq n^{-1/(3+d)}$, we have that
	\begin{equation*}
	\Ebb(R_{s,n}(f)) = O(L^2)
	\end{equation*} 
\end{lemma}

\section{\textcolor{red}{Bounding the Roughness Functional in Expectation}}

\textcolor{red}{WARNING:} The following Lemma is not proved, but it is the result I am trying to prove. After the Lemma is stated, I go into some detail about how I intend to prove it.

For an integer $\ell > 0$, we will say $K$ is an order-$\ell$ kernel
\begin{equation*}
\int K(z) = 1 ~~\textrm{and}~~\int z^j K(z) = 0,~~\textrm{for $j = 1,\ldots,\ell$}
\end{equation*}
and further $K$ is $\ell$ times continuously differentiable with $K^{j}(z) = O(z^j K(z))$ for $j = 1,\ldots,s$.

\begin{lemma}
	\label{lem:roughness_functional_expectation}
	Suppose $f \in C^{s}(L)$ for some $s,L > 0$, and further suppose $p \in C^{\ell}(L)$ for $\ell = \floor{s}$ the largest integer strictly less than $s$. For any order-$\ell$ kernel $K$ and any $r(n) \geq n^{-1/(4+d)}$, we have that
	\begin{equation*}
	\Ebb(R_{s,n}(f)) = O(r^{2s})
	\end{equation*} 
\end{lemma}
When $s \geq 3$ the graph Laplacian approximates higher-order derivatives by composing difference operators. The crudeness of this approximation means that in Lemma~\ref{lem:roughness_functional_expectation} we allow for a slower scaling of $r(n)$ than we did Lemma~\ref{lem:2nd_order_roughness_functional}. At a high level, however, we prove Lemma~\ref{lem:roughness_functional_expectation} in a similar fashion to Lemma~\ref{lem:2nd_order_roughness_functional}. We first break the sum into cases according to the number of unique indices, then bound the expectation case by case. To simplify our statements, for index vectors $k,\ell \in [n]^q$ we will use the following notation:
\begin{equation*}
\textrm{Unq}(k \setminus \ell) = \set{x \in k: x \not\in \ell, \textrm{$x$ appears in $k$ exactly once,}}~~ k \bigcup \ell = \set{x: x \in k ~\textrm{or}~ x \in \ell}
\end{equation*}
Thus $\textrm{Unq}(k \setminus \ell)$ measures the number of indices in $k$ which appear exactly once in $k$ and never in $\ell$, and $k \bigcup \ell$ measures the total number of distinct indices.

The following Lemma will be the workhorse which supplies a sufficient bound in all cases.
\begin{lemma}
	\label{lem:expected_difference_operators}
	Let $q = s/2$ when $s$ is even, and $q = (s - 1)/2$ when $s$ is odd. Under the same conditions as Lemma~\ref{lem:roughness_functional_expectation}, for any indices $k = (k_1,\ldots,k_q)$ and $\ell = (\ell_1,\ldots,\ell_q)$, we have
	\begin{equation}
	\Ebb(D_kf(x_i) D_kf(x_i)) =
	\begin{cases*}
	O(r^{2} r^{d(\abs{k \cup \ell} - 2q)}), & ~~\textrm{if $\Unq(k\setminus\ell) = \emptyset$ and $\Unq(\ell\setminus k) = \emptyset$}~ 
	\\
	O(r^{2q + 1} r^{\abs{k \cup \ell}} r^{d(\abs{k \cup \ell} - 2q)}), & ~~\textrm{if $\Unq(k\setminus\ell) \neq \emptyset$ or $\Unq(\ell\setminus k) \neq \emptyset$} \\
	O(r^{4q} r^{d(\abs{k \cup \ell} - 2q)}), & ~~\textrm{if $\Unq(k\setminus\ell) \neq \emptyset$ and $\Unq(\ell\setminus k) \neq \emptyset$}
	\end{cases*}
	\end{equation}
\end{lemma}

\section{Proofs}
\subsection{Proof of Lemma~\ref{lem:2nd_order_roughness_functional}}
We prove Lemma~\ref{lem:2nd_order_roughness_functional} in the case when $d = 1$.

\textcolor{red}{TODO:} Extend the proof to all values of $d$. It's not hard, just requires you to replace derivatives by partial derivatives.

Separating the squared sum in~\eqref{eqn:roughness_functional_representation_even} into diagonal and off-diagonal terms, we obtain
\begin{align*}
R_{2,n}(f) & = \frac{1}{n}\sum_{i = 1}^{n}\left(\frac{1}{nr^2}\sum_{j = 1}^{n} D_jf(x_i)\right)^2 \\
& = \frac{1}{n^3r^4}\sum_{i = 1}^{n} \sum_{j = 1}^{n}\underbrace{(D_jf(x_i))^2}_{V_1} + \frac{1}{n^3r^4}\sum_{i = 1}^{n} \sum_{j = 1} \sum_{k \neq j}\underbrace{(D_jf(x_i))(D_kf(x_i))}_{U_1}
\end{align*}

The expectation of $U_1$ will be of the right order of magnitude thanks to a cancellation of the first-order term in a Taylor expansion. Noting that the kernel $K$ satisfies
\begin{equation*}
\int z K(z) \,dz = 0
\end{equation*}
we therefore derive that for any $x \in \mathcal{X}$,
\begin{align*}
\mathbb{E}(D_jf(x)) & = \frac{1}{r}\int ((y - x)f'(x) + O(L(y-x)^2) K(y - x) p(y) \,dy \tag{by $f \in C^2(L)$} \\
& = \frac{1}{r}\int(y - x)K(y - x)f'(x)(p(x) + O(L(y - x))) \,dy + O(Lr^2)  \tag{by $p \in C^1(L)$}\\
& = \underbrace{\frac{f'(x) p(x)}{r} \int z K(z) \,dz}_{= 0} + O(Lr^2) = O(Lr^2).
\end{align*}
As a result, we have
\begin{equation*}
\frac{1}{n^3r^4}\sum_{i = 1}^{n} \sum_{j = 1} \sum_{k \neq j} \Ebb\left[(D_jf(x_i))(D_kf(x_i))\right] = \frac{1}{n^3r^4}\sum_{i = 1}^{n} \sum_{j = 1} \sum_{k \neq j} \Ebb\left[ \Ebb(D_kf(x)\mid x_i = x)^2\right] = O(L^2).
\end{equation*}

The expectation of the diagonal terms may be much larger, but on the other hand there are many fewer of them. Specifically, we compute
\begin{align*}
\Ebb\bigl((D_kf(x_i))^2\bigr) & = \Ebb\biggl(O\left(L^2(x_k - x_i)^2\right)\cdot K_r(x_k,x)^2\biggr) \\
& = O(L^2r^2)\cdot\Ebb(K_r(x_k,x_i)^2) = O(L^2 r)
\end{align*}
and therefore
\begin{equation*}
\frac{1}{n^3r^4}\sum_{i = 1}^{n} \sum_{j = 1}^{n} \Ebb\bigl((D_jf(x_i))^2\bigr) = \frac{1}{n} O(L^2 r^{-3}) = O(L^2)
\end{equation*}
where the second equality follows since by assumption when $d = 1$, $n^{-1} = O(r^{3})$.

\section{Proof of Lemma~\ref{lem:3rd_order_roughness_functional}}

\textcolor{red}{TODO}: Copy proof of Lemma~\ref{lem:3rd_order_roughness_functional} from handwritten notes.

\section{Proof of Lemma~\ref{lem:expected_difference_operators}}

\textcolor{red}{TODO}: Prove Lemma~\ref{lem:expected_difference_operators}.

\section{Auxiliary Results.}
\begin{lemma}
	\label{lem:difference_operator_taylor}
	When $f \in C^s(L)$, for any $k \in [n]$ 
	\begin{equation*}
	D_kf(x) = \left(\sum_{r = 1}^{s - 1} (x_k - x)^r f^{(r)}(x) + O((x_k - x)^{s})\right) \cdot K_r(x_k,x)
	\end{equation*}
\end{lemma}

\end{document}

### 2020-01-09 notes/weekly_notes/1.3.20.NotesForWeek.tex
added weekly notes for break

--- 2019-12-10 notes/graph_testing.tex
@@
 \newcommand{\wt}[1]{\widetilde{#1}}
+\newcommand{\ol}[1]{\overline{#1}}
 \newcommand{\spec}{\mathrm{spec}}
+\newcommand{\Unq}{\mathrm{Unq}}
 
@@
 \begin{document}
-\title{Graph Testing}
+\title{Notes for Week 12/30/19 - 1/3/19}
 \author{Alden Green}
@@
 
-This document details the current status of the graph testing project.
-We divide by section based on the testing problem under consideration.
-At the end of each section, we list some areas we are interested in investigating.
-All proofs are left until the end.
-
-Throughout, let $X = \{x_1,
-\ldots, x_n\}$ be a sample drawn i.i.d. from a distribution $P$ on $\Rd$,
-with density~$p$.  For a radius $r > 0$, we define $G_{n,r}=(V,E)$ to be the
-\emph{$r$-neighborhood graph} of $X$, an unweighted, undirected graph with
-vertices $V=X$, and an edge $(x_i,x_j) \in E$ if and only if $K_r(x_i,x_j) = \norm{x_i -x_j} \leq r$, where $\norm{\cdot}$ is the Euclidean norm. We denote by $A \in
-\Reals^{n \times n}$ the adjacency matrix, with entries $A_{uv} = 1$ if
-$(u,v) \in E$ and $0$ otherwise.  We also denote by $D$ the diagonal degree
-matrix, with entries $D_{uu} := \sum_{v \in V} A_{uv}$. The graph Laplacian is $L = D - A$, and we write its spectral decomposition as $L = V S V^T$.
-
-\section{Regression goodness-of-fit testing with random design.}
-
-Let $P$ be a distribution with density $p$ supported on $\mathcal{X} \subseteq \Reals^d$. Suppose we observe random design points $X = \set{x_1,\ldots,x_n} \sim P$, and additionally responses
+Suppose we observe independent samples $X = \set{x_1,\ldots,x_n}$ from a distribution $P$ with density $p$ supported on $\mathcal{X} = [0,1]^d$.
+Let $G_{n,r}$ be the random geometric graph of radius $r$ formed over the samples $X$, that is, the neighborhood graph with kernel $K(x) = \1(\norm{x} \leq r)$, and let $L$ denote the corresponding graph Laplacian matrix.
+In this week's notes, we will show that as long as $r$ is chosen to be sufficiently small, the eigenvalues of $L$ we care about scale at an appropriate rate.
+More precisely, we will prove the following Lemma.
+\begin{lemma}
+	\label{lem:higher_order_eigenvalue_tail_bound}
+	Let $\kappa = n^{2d/(4s + d)}$. Suppose $p$ is uniformly bounded away from $0$ and $\infty$ on its support,
+	\begin{equation*}
+	0 < p_{\min} < p(x) < p_{\max} < \infty,~~\textrm{for all $x \in \Xset$.}
+	\end{equation*}
+	Then there exists a constant $c$ such that for any $r \leq n^{-4/((2+d)(4s + d))}$, the eigenvalue $\lambda_{\kappa}$ is lower bounded,
 	\begin{equation}
-\label{eqn:regression_known_variance}
-y_i = f(x_i) + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
+	\label{eqn:higher_order_eigenvalue_tail_bound}
+	\lambda_{\kappa} \geq c\kappa^{2/d}nr^{2 + d}
 	\end{equation}
-
-We wish to distinguish
-\begin{equation*}
-\mathbf{H}_0: f = f_0 := 0 \quad \textrm{vs} \quad \mathbf{H}_{\textrm{a}}: f \neq f_0
-\end{equation*}
-We will evaluate our performance using worst-case risk: for a given function class $\mathcal{H}$ and test function $\phi: \Reals^n \to \set{0,1}$, let
-\begin{equation*}
-\mathcal{R}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, f \neq f_0} \Ebb_f(1 - \phi).
-\end{equation*}
-The worst-case risk may be quite close to $1$ unless we require some separation between null and alternative spaces. A more realistic measure of performance is therefore
-\begin{equation*}
-\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, \norm{f - f_0}_2 \geq \epsilon} \Ebb_f(1 - \phi).
-\end{equation*}
-
-\subsection{Test Statistics.}
-We list the test statistics we use for the regression testing problem.
-
-\paragraph{Eigenvector projection test statistic.}
-Let $VSV^T$ be the spectral decomposition of the Laplacian matrix $L$ of the neighborhood graph $G_{n,r}$.
-To test whether $f = f_0$, we propose the following \emph{eigenvector projection} test statistic:
-\begin{equation}
-\label{eqn:graph_spectral_projections}
-T_{\mathrm{spec}} := \frac{1}{n} \sum_{k = 0}^{\kappa} \left(\sum_{i = 1} v_i y_i\right)^2
-\end{equation}
-
-\subsection{Current Results.}
-
-In Theorem~\ref{thm:sobolev_testing_rate} we show that under some typical regularity conditions on $P$, the test $\phi_{\textrm{spec}} := \1\{T_{\mathrm{spec}} \geq \tau\}$ is, up to log factors, a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};R)$.
-To conveniently state our results we introduce the notation
-\begin{equation*}
-h(a,d) = a(d+2) + (1 + 2/d),~~\textrm{for $a > 0$}
-\end{equation*}
-
-\begin{theorem}
-	\label{thm:sobolev_testing_rate}
-	Let $b \geq 1$ and $a > 0$ be fixed constants, and let $d < 4$.
-Suppose that $\Pbb$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p$ bounded above and below by constants, i.e
-	\begin{equation*}
-	0 < p_{\min} < p(x) < p_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
-	\end{equation*}
-	Then the following statement holds: if the test $\phi_{\spec}$ is performed with parameter choices 
-	\begin{equation*}
-	r = \log^a n \cdot \left(\frac{\log n}{n}\right)^{1/d}, ~\kappa = n^{2d/(4 + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
-	\end{equation*}
-	then there exists a constant $c$ which may depend on $d,R,p_{\max},b$ and $a$ but is independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
-	\begin{equation}
-	\label{eqn:sobolev_testing_rate}
-	\epsilon^2 \geq c \cdot b \cdot n^{-4/(4 + d)} (\log n)^{h(a,d)}
-	\end{equation}
-	the worst-case risk is upper bounded
-	\begin{equation}
-	\label{eqn:sobolev_testing_rate_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; \mathcal{W}^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}.
-	\end{equation}
-\end{theorem}
-The testing rate~\eqref{eqn:sobolev_testing_rate} matches the minimax critical radius up to a factor of $\log^{h(a,d)}n$.
-
-When $d \geq 4$ the compact embedding
-\begin{equation*}
-W_d^{1,2}(\mathcal{X}) \subseteq \mathcal{L}_d^4(\mathcal{X}) 
-\end{equation*}
-does not hold (as it does when $d < 4$). However, if we directly assume $f \in \mathcal{L}_d^4(\mathcal{X};R)$ (regardless of $d$) we obtain the following result.
-\begin{proposition}
-	\label{prop:L4_testing_rate}
-	If $f \in \mathcal{L}_d^4(\mathcal{X};R)$, there exists a constant $c$ such that if
-	\begin{equation}
-	\label{eqn:L4_testing_rate}
-	\epsilon^2 > b \cdot n^{-1/2}
-	\end{equation}
-	then the test
-	\begin{equation*}
-	\phi_{\mathrm{mean}} = \1\{\frac{1}{n}\sum_{i = 1}^{n} y_i^2 \geq 1\}
-	\end{equation*}
-	has worst-case risk
-	\begin{equation*}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}. 
-	\end{equation*}
-\end{proposition}
-
-Note that when $d < 4$ the rate~\eqref{eqn:sobolev_testing_rate} is sharper than \eqref{eqn:L4_testing_rate}.
-
-\subsection{Areas to Investigate}
-
-\begin{enumerate}[(i)]
-	\item When $d \geq 4$, is it true that there does not exist a uniformly consistent
-	test over the Sobolev ball $W_d^{1,2}(\mathcal{X};R)$?
-	\item Does the testing rate of $\phi_{\mathrm{spec}}$ over Sobolev spaces $W_d^{s,2}(\mathcal{X};R)$ match the minimax rate, for an appropriate choice of kernel $K$?
-	\item Assume the distribution $P$ is supported on a manifold $\mathcal{M}$ of intrinsic dimension $s < d$.
-Does $\mathrm{\phi_{\mathrm{spec}}}$ display adaptivity to the intrinsic dimension of $\mathcal{M}$?
-	\item Assume that $f$ belongs to the Holder space $C_d^s(\mathcal{X})$.
-Moreover, suppose that instead of observing ${y_i}$ according to the regression testing model \eqref{eqn:regression_known_variance}, we observe
-	\begin{equation*}
-	y_i = f(x_i) + \sigma \varepsilon_i, ~\varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
-	\end{equation*}
-	where $\sigma > 0$ is unknown.
-When $d \geq 4$, what are the minimax regression testing rates over $C_d^1(\mathcal{X};R)$?
-Is the test $\phi_{\mathrm{spec}}$ minimax optimal, when the tuning parameters $r$ and $\kappa$ are appropriately chosen?
-\end{enumerate}
-
-\section{Two-sample density testing.}
-In the two-sample density testing problem, we observe independent samples $Z = z_1,\ldots,z_N \sim P$ and $Y = y_1,\ldots,y_M \sim Q$, where $P$ and $Q$ are distributions over $\Reals^d$ with densities $p$ and $q$, respectively, and $N \sim \textrm{Bin}(n,1/2)$.
-Our goal is to distinguish the hypotheses
-\begin{equation*}
-\mathbf{H}_0: P = Q \quad \textrm{vs} \quad \mathbf{H}_{\textrm{a}}: P \neq Q
-\end{equation*}
-and we again evaluate our performance using worst-case risk; letting $\phi:\Reals^{N + M} \to \{0,1\}$, 
-\begin{equation*}
-\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \inf_{p \in \mathcal{H}}\Ebb_{p,p}(\phi) + \sup_{\substack{p,q \in \mathcal{H} \\ \norm{p - q}_{\Leb^2} \geq \epsilon}} \Ebb_{p,q}(1 - \phi).
-\end{equation*}
-
-\subsection{Test statistics.}
-
-We suggest several two-sample test statistics.
-
-\paragraph{Eigenvector projection test statistic.}
-
-It is straightforward to adapt the test statistic $T_{\mathrm{spec}}$ to the two-sample testing problem.
-Concatenate the samples $Z$ and $Y$ in $X = (z_1,\ldots,z_N,y_1,\ldots,y_M)$, and define $T_{\mathrm{spec}}^{(2)}$ to be
-\begin{equation}
-\label{eqn:graph_spectral_projections_2}
-T_{\mathrm{spec}}^{(2)} := \frac{1}{n} \sum_{k = 0}^{\kappa} \left(\sum_{i = 1}^{n} v_i a_i\right)^2, ~~\textrm{where}~~ a = (\underbrace{N^{-1},\ldots,{N^{-1}}}_{\textrm{length } N},\underbrace{-M^{-1},\ldots,-M^{-1}}_{\textrm{length } M})
-\end{equation}
-
-For convenience, we state our following two test statistics with respect to the empirical norm $\norm{\theta}_n = n^{-1/2}\norm{\theta}_2$ for $\theta \in \Reals^n$.
-They will each depend on a tuning parameter $\lambda > 0$.
-\paragraph{Graph Sobolev IPM.}
-Letting $C_n := nr^{(d + 2)/2}$ and
-\begin{equation*}
-\Theta_{1,2} := \{\theta \in \Reals^n:~ C_n^{-1} \norm{B\theta}_2 \leq 1\} 
-\end{equation*}
-we define the \emph{graph Sobolev IPM} to be
-\begin{equation}
-\label{eqn:sobolev_IPM}
-T_{\textrm{sob}} := \sup_{\substack{\theta \in \Theta_{1,2} \\ \lambda \norm{\theta}_n \leq 1}} \abs{\frac{1}{n}\sum_{i = 1}^{n} a_i \theta_i}. 
-\end{equation}
-
-\paragraph{Graph Total Variation IPM.}
-Letting $C_n' := n^{2}r^{(d + 1)}$ and 
-\begin{equation*}
-\Theta_{1,1} := \{\theta \in \Reals^n:~ (C_n')^{-1} \norm{B\theta}_1 \leq 1\}
-\end{equation*}
-we define the \emph{graph Total Variation} IPM to be
-\begin{equation}
-\label{eqn:total_variation_IPM}
-T_{\mathrm{TV}} := \sup_{\substack{\theta \in \Theta_{1,1}, \\ \lambda \norm{\theta}_n \leq 1} } \abs{\frac{1}{n}\sum_{i = 1}^{n} a_i \theta_i}, \quad
-\end{equation}
-
-\subsection{Current results.}
-
-\textcolor{red}{WARNING:} While I have proved all the parts I believe are required for the below theorem, I haven't actually put together the pieces yet.
-
-In Theorem~\ref{thm:twosample_sobolev_testing_rate} we show that under some typical regularity conditions on $P$, the test $\phi_{\textrm{spec}}^{(2)} := \1\{T_{\mathrm{spec}}^{(2)} \geq \tau\}$ is, up to log factors, a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};R)$ when $d = 1$.
-
-\begin{theorem}
-	\label{thm:twosample_sobolev_testing_rate}
-	Let $b \geq 1$ and $a > 0$ be fixed constants, and let $d = 1$.
-Suppose that $P$ and $Q$ are absolutely continuous probability measures over $\mathcal{X} = [0,1]$ with density functions $p$ and $q$ bounded above and below by constants, i.e
-	\begin{equation*}
-	0 < p_{\min},q_{\min} < p(x),q(x) < p_{\max},q_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
-	\end{equation*}
-	Then the following statement holds: if the test $\phi_{\spec}^{(2)}$ is performed with parameter choices 
-	\begin{equation*}
-	r = \log^a n \cdot \left(\frac{\log n}{n}\right), ~\kappa = n^{2/5}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
-	\end{equation*}
-	then there exists a constant $c$ which may depend on $R,p_{\max},q_{\max},b$ and $a$ but is independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
-	\begin{equation}
-	\label{eqn:twosample_sobolev_testing_rate}
-	\epsilon^2 \geq c \cdot b \cdot n^{-4/5} (\log n)^{h(a,1)}
-	\end{equation}
-	the worst-case risk is upper bounded
-	\begin{equation}
-	\label{eqn:twosample_sobolev_testing_rate_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; \mathcal{W}^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}.
-	\end{equation}
-\end{theorem}
-
-We prove Theorem~\ref{thm:twosample_sobolev_testing_rate} by relating the density testing problem to a regression testing problem with a certain type of structured noise, and then proceeding along similar lines to the proof of Theorem~\ref{thm:sobolev_testing_rate}.
-To pursue this strategy, we require the eigenvectors to satisfy a certain type of incoherence condition; this is in constrast to the regression testing problem with known variance, where we did not require the eigenvectors to be smooth in any sense.
-
-\subsection{Areas to Investigate}
-
-\begin{enumerate}[(i)]
-	\item When $1 < d < 4$, is the test $\phi_{\spec}^{(2)}$ minimax optimal?
-	\item Are either or both of the test statistics \eqref{eqn:graph_spectral_projections_2}-\eqref{eqn:sobolev_IPM} minimax optimal over $C_d^1(\mathcal{X};R)$ for all values of $d$?
-	\item Are either or both of the test statistics \eqref{eqn:graph_spectral_projections_2}-\eqref{eqn:sobolev_IPM} minimax optimal over $W_d^{1,2}(\mathcal{X};R)$ when $d \leq 4$?
-	\item Is the test statistic \eqref{eqn:graph_spectral_projections_2}, computed over a graph with suitable choice of kernel $K$, minimax optimal over higher order derivative classes $C_d^s$ and $W^{s,2}$?
-	\item Modify the test statistic \eqref{eqn:sobolev_IPM} by replacing the function class $\Theta_{1,2}$ with
-	\begin{equation}
-	\Theta_{s,2} := \{\theta \in \Reals^n:~ C_n^{-1} \norm{B^{(s)}\theta}_2 \leq 1\} 
-	\end{equation}
-	Is the modified test statistic, computed over a graph with suitable choice of kernel $K$, minimax optimal over higher order derivative classes $C_d^s$ and $W^{s,2}$?
-	\item What is the minimax testing rate over $BV_d^{1}(\mathcal{X};R)$? Does it exhibit a phase transition analogous to the minimax estimation rate over bounded variation spaces?
-	\item Is the test statistic \eqref{eqn:total_variation_IPM} minimax optimal over $BV_d^{1}(\mathcal{X};R)$?
-	\item Is the test statistic \eqref{eqn:total_variation_IPM} minimax optimal over Sobolev and Holder function classes?
-\end{enumerate}
-
-\section{Definitions}
-
-Here we collect definitions of some common function spaces and graph operators.
-
-\subsection{Function Spaces}
-
-\paragraph{Lebesgue spaces.}
-
-We say a Borel measurable function $f: \mathcal{X} \to \Reals$ is in the space $\mathcal{L}^p(\mathcal{X})$ for $1 \leq p < \infty$ if 
-$$\norm{f}_{\mathcal{L}^p(\mathcal{X})} := \int_{\mathcal{X}} \abs{f(x)}^p \,dx < \infty$$
-and we let 
-\begin{equation*}
-\mathcal{L}^p(\mathcal{X};R) = \set{f \in \mathcal{L}^p(\mathcal{X}): \norm{f}_{\mathcal{L}^p(\mathcal{X})} < R}
-\end{equation*}
-be a ball in the Lebesgue space.
-
-\paragraph{Holder spaces.}
-
-For a given $s > 0$, the $s$th Holder norm is given by
-\begin{equation*}
-\norm{f}_{C_d^{s}(\mathcal{X})} := \sum_{\abs{\alpha} \leq s} \norm{D^{\alpha}f}_{\infty} + \sum_{\abs{\alpha} = s} \sup_{x,y \in \mathcal{X}} \frac{\abs{D^{\alpha}f(y) - D^{\alpha}f(x)}}{\norm{x - y}_2}
-\end{equation*}
-and the $s$th Holder space $C_d^{s}(\mathcal{X})$ consists of all functions which are $s$ times continuously differentiable with finite $s$ Holder norm.
-Denote the Holder  ball by $C_d^{s}(\mathcal{X},R) = \set{f \in C_d^{s}(\mathcal{X}): \norm{f}_{C_d^{s}(\mathcal{X})} \leq R}$.
-
-\paragraph{Sobolev spaces.}
-
-For a given $s > 0$, the Sobolev space $W_d^{s,2}(\mathcal{X})$ consists of all functions $f \in \mathcal{L}^2(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^2(\mathcal{X})$.
-The Sobolev $\{s,2\}$ norm is then
-\begin{equation*}
-\norm{f}_{W_d^{s,2}(\mathcal{X})}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^2 \,dx
-\end{equation*}
-and for a given $L > 0$, the corresponding ball is $W_d^{s,2}(\Xset; L) = \set{f: \norm{f}_{W^{s,2}(\Xset)} \leq L}$.
-
-\paragraph{Bounded Variation spaces.}
-
-For a function $f \in L^1(\mathcal{X})$ the \emph{total variation} semi-norm of $f$ is
-\begin{equation*}
-TV(f;\mathcal{X}) := \sup \left\{ \int_{\mathcal{X}} f \, \Xsetive \, \psi \,dx : \psi \in C_c^1(\mathcal{X}; \Reals^d), \abs{\psi} \leq 1 \right\};
-\end{equation*}
-and we write $BV_d(\mathcal{X})$ for the subset of functions $f \in L^1(\mathcal{X})$ which have bounded norm
-\begin{equation*}
-\norm{f}_{BV_d(\mathcal{X})} := \norm{f}_{\infty} + TV(f;\mathcal{X}).
-\end{equation*}
-For a given $R > 0$, the corresponding ball is $BV_d^{1}(\mathcal{X};R) = \set{f: \norm{f}_{BV_d(\mathcal{X})} \leq R}$.
-
-\subsection{Graph Operators.}
-Let $s \geq 1$ be an integer. The $s$th-order difference operator on $G_{n,r}$, denoted $B^{(s)}$, is defined by
-\begin{equation*}
-B^{(s)} :=
-\begin{cases}
-L^{s/2},& ~~ s \textrm{ even} \\
-BL^{(s - 1)/2},& ~~ s \textrm{ odd.}
-\end{cases}
-\end{equation*}
+	with probability $1 - o(n^{-1})$ as $n \to \infty$.
+\end{lemma}
+Lemma~\ref{lem:higher_order_eigenvalue_tail_bound} establishes that eigenvalues of the graph Laplacian, suitably normalized, scale comparably to the eigenvalues of the continuum Laplacian operator $\mathcal{D}f := \textrm{div}(\nabla f)$.
+Along with an upper bound on the graph Sobolev semi-norm $R_{s,n}(f)$ (established in the \textit{12.20.19 Notes}), this allows us to argue that if $f$ is a smooth function with large $\mathcal{L}_2$ norm, then the projection of a function $f$ onto the eigenvectors $v_1,\ldots,v_{\kappa}$ is still quite large (measured in empirical norm); or, equivalently, the bias induced by projection onto these eigenvectors is small.
 
@@
 
-\subsection{Proof of Theorem~\ref{thm:sobolev_testing_rate}}
+\subsection{Proof of Lemma~\ref{lem:higher_order_eigenvalue_tail_bound}}
 
-To prove Theorem~\ref{thm:sobolev_testing_rate}, we will show that when
-\begin{equation}
-\label{eqn:critical_radius}
-\epsilon^2 \geq c \cdot b \cdot (\log n)^{h(\beta,d)} \cdot n^{-4/(4 + d)}
-\end{equation}
-and we choose tuning parameters as specified in Theorem~\ref{thm:sobolev_testing_rate}, the test $\phi_{\mathrm{spec}}$ satisfies
-\begin{equation}
-\mathcal{R}_{\epsilon}(\phi_{\spec}; W^{1,2}(\mathcal{X};R)) = \Ebb_{f_0}(\phi_{\spec}) + \sup_{\substack{f \in W^{1,2}(\mathcal{X};R) \\ \norm{f}_{\Leb^2(\mathcal{X})} > \epsilon}} \Ebb_{f}(1 - \phi_{\spec}) \leq \left(\frac{1}{b^2} + \frac{3}{b}\right) + o(1).
-\end{equation}
+We prove Lemma~\ref{lem:higher_order_eigenvalue_tail_bound} by comparing $G_{n,r}$ to the tensor product of a $d$-dimensional lattice and a complete graph.
+The latter is a highly structured graph with known eigenvalues, which as we will see are sufficiently lower bounded for our purposes.
 
-Our analysis will proceed according to the following steps.
-\begin{enumerate}
-	\item We upper bound the testing error of our eigenvector projection test statistic when $Y$ are viewed as random responses defined over the vertices of a fixed graph $G$.
-This upper bound will hold whenever certain functionals on the graph $G$ are themselves bounded.
-	\item We analyze the behavior of these functionals with respect to the random graph $G_{n,r}$, and bound them with high probability.
-	\item We condition on $X \subseteq \mathcal{E}$, where $\mathcal{E}$ is a high-probability set over which the relevant functionals on $G_{n,r}$ are bounded.
-We conclude that the upper bound on testing error derived in our first step holds whenever $X \subseteq \mathcal{E}$.
-\end{enumerate}
+Let $\wt{r} = r/(3(\sqrt{d} + 1)), M = (1/\wt{r})^d, N = n\wt{r}^d$. Assume without loss of generality that $M$ and $N$ are integers. Additionally, for $t = n^{1/d}$ and $m = M^{1/d}$ let 
+\begin{equation*}
+\overline{X} = \set{\frac{1}{t}(k_1,\ldots,k_d): k \in [t]^d},~~ \overline{Z} = \set{\frac{1}{m}(j_1,\ldots,j_d): j \in [m]^d}.
+\end{equation*}
+For a given $\overline{z}_j \in \overline{Z}$, we write $Q(z_j) = m^{-1}[j_1 - 1,j_1] \times \cdots \times m^{-1}[j_d - 1,j_d]$ for the cube of side length $1/m$ with $z_j$ at one corner.
 
-\subsubsection{Step 1: Testing error on a fixed graph}
-
-Let $G = (V,E)$ be a graph over vertices $V = \set{v_1,\ldots,v_n}$, and let $\beta = (\beta_1,\ldots,\beta_n) \in \Reals^n$ be a signal over the vertices $V$.
-We observe responses $Y = (y_1,\ldots,y_n)$ according to the model
+Consider the graph $H = (\overline{X}, E_H)$, where $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$ if
 \begin{equation*}
-y_i = \beta_i + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
+\textrm{there exists}~\ol{z}_i, \ol{z}_j \in \ol{Z}~\textrm{such that}~\ol{x}_k \in Q(\ol{z}_i),~ \ol{x}_\ell \in Q(\ol{z}_j),~\textrm{and}~\norm{i - j}_1 \leq 1.
 \end{equation*}
-Letting $L = VSV^T$ be the spectral decomposition of the Laplacian $L$ of $G$, our graph spectral test statistic is 
+On the one hand $H \cong \ol{G}^M_d \otimes K_N$ where $\ol{G}^M_d$ is the $d$-dimensional lattice on $M$ nodes, and $K_N$ is the complete graph on $N$ nodes.
+On the other hand, we now show that with high probability $G_{n,r} \succeq H$.
+If $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$, then there exist $\ol{z}_i, \ol{z}_j$ such that
 \begin{equation*}
-T_{\spec} = \frac{1}{n}\sum_{k = 1}^{\kappa} \left(\sum_{i = 1}^{n} v_i y_i\right)^2
-\end{equation*}
-where $\kappa$ is a tuning parameter. The resulting test we will use is
-\begin{equation*}
-\phi_{\spec} = \1\{T_{\spec} \geq \frac{\kappa}{n} + t(b)\},~~\textrm{where}~ t(b) = b\sqrt{\frac{2\kappa}{n^2}}~~\textrm{for $b \geq 1$.}
+\norm{\ol{x}_k - \ol{x}_{\ell}}_2 \leq m^{-1} + \norm{\ol{x}_k - \ol{z}_{i}}_2 + \norm{\ol{x}_{\ell} - \ol{z}_{j}}_2 \leq \wt{r}(1 + \sqrt{d}) = r/3.
 \end{equation*}
 
-Let $S_2(\beta;G)$ be a measure of smoothness the signal $\beta$ displays over the graph $G$, given by
+By Theorem 1.1 of \textcolor{red}{Garcia Trillos and Slepcev}, with probability at least $1 - n^{-1}$ there exists a bijection $\pi: \overline{X} \to X$ such that
+\begin{equation}
+\label{eqn:transport_distance}
+\max_{k \in [t]^d} \abs{\overline{x}_k - \pi(\overline{x}_k)} \leq c \left(\frac{\log n}{n}\right)^{1/d}
+\end{equation}
+Assuming~\eqref{eqn:transport_distance} holds, if $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$, then for sufficiently large $n$
 \begin{equation*}
-S_2(\beta;G) = \sum_{i = 1}^{n} \sum_{j = 1}^{n} (\beta_i - \beta_j)^2 \1((v_i,v_j) \in E).
+\norm{\pi(\ol{x}_k) - \pi(\ol{x}_\ell)}_2 \leq 2 c \left(\frac{\log n}{n}\right)^{1/d} + \frac{r}{3} \leq r,
 \end{equation*}
-In Lemma~, we upper bound the Type I and Type II error of the test $\phi_{\spec}$.
-Our bound on the Type II error will be stated as a function of $S_2(\beta;G)$ as well as the $\kappa$th eigenvalue $s_{\kappa}$.
+implying that $(\pi(\ol{x}_k), \pi(\ol{x}_{\ell})) \in E$. Therefore, $G_{n,r} \succeq \ol{G}^M_d \otimes K_N$ whenever~\eqref{eqn:transport_distance} holds.
 
-\begin{lemma}
-	\label{lem:fixed_graph_testing}
-	Let $1 \leq \kappa \leq n$ be an integer.
-	\begin{enumerate}
-		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}$ is upper bounded
-		\begin{equation}
-		\label{eqn:graph_spectral_type_I_error}
-		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{1}{b^2}.
-		\end{equation}
-		\item \textbf{Type II error:} For any $b$ and $\beta$ such that
-		\begin{equation}
-		\label{eqn:fixed_graph_testing_critical_radius}
-		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_2(\beta;G)}{ns_{\kappa}}
-		\end{equation}
-		the Type II error of $\phi_{\beta}$ is upper bounded,
-		\begin{equation}
-		\label{eqn:graph_spectral_type_II_error}
-		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{3}{b}.
-		\end{equation}
-	\end{enumerate}
-\end{lemma}
-
-The first term on the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius} grows with $\kappa$, while the second term shrinks.
-These can be thought of as the variance and squared bias terms, respectively, of the error.
-
-\paragraph{Proof of Lemma~\ref{lem:fixed_graph_testing}.}
-
-To prove Lemma~\ref{lem:fixed_graph_testing} we will first compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}$, and then use Chebyshev's inequality to show \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.
-
-\vspace{.2 in}
-
-\textit{Mean of $T_{\mathrm{spec}}$:} Using the notation $\dotp{v}{w} = \sum_{i = 1}^{n} v_iw_i$, we have
+The eigenvalues of lattices and complete graphs are known to satisfy, respectively
+\begin{equation*}
+\lambda_k(\ol{G}^{M}_d) \geq \frac{k^{2/d}}{M^{2/d}}~\textrm{for $k = 0,\ldots,M - 1$},~~ \textrm{and}~\lambda_{j}(K_N) \geq N\1\{j > 0\}~\textrm{for $j = 0,\ldots,N-1$.}
+\end{equation*}
+and by standard facts regarding the eigenvalues of tensor product graphs (\textcolor{red}{reference}), we have that the spectrum $\Lambda(H)$ satisfies
+\begin{equation*}
+\Lambda(H) = \set{N\lambda_k(\ol{G}^{M}_d) + M\lambda_j(K_N): \textrm{for $k = 0,\ldots,M - 1$ and $j = 0,\ldots,N-1$}}
+\end{equation*}
+For all $j = 1,\ldots,N-1$, we have that $M\lambda_j(K_N) = MN = n$. Therefore,
 \begin{align*}
-\Ebb(T_{\spec}) & = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{\beta}{v_k}^2 + \Ebb\bigl( \dotp{\varepsilon}{v_k}^2 + 2 \dotp{\varepsilon}{v_k} \dotp{\beta}{v_k}\bigr)\right) \\
-& = \frac{\kappa}{n} + \frac{1}{n}\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2.
+\lambda_{\kappa}(H) & \geq \{n \wedge N\lambda_{\kappa}(\ol{G}^{M}_d)\} \\
+& \geq \{n \wedge n\wt{r}^d\frac{\kappa^{2/d}}{M^{2/d}}\} \\
+& \geq \{n \wedge (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d}\} \\
+& \geq (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d},
 \end{align*}
-When $\beta = 0$, this equals $\kappa/n$. Otherwise, we have the following lower bound:
-\begin{align*}
-\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2 & = \norm{\beta}_2^2 - \sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \\
-& \geq \norm{\beta}_2^2 - \frac{1}{s_{\kappa}}\sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 s_k \\
-& \geq \norm{\beta}_2^2 - \frac{S_2(\beta;G)}{s_{\kappa}},
-\end{align*}
-and therefore $\Ebb(T_{\spec}) \geq \kappa/n + n^{-1}(\norm{\beta}_2^2 - S_2(\beta;G)/s_{\kappa})$.
-
-\vspace{.2 in}
-
-\textit{Variance of $T_{\mathrm{spec}}$:}
-We write $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvectors $v_1,\ldots,v_{\kappa}$ as columns. Consequently,
-\begin{align}
-\Var(T_{\spec}) & = \frac{1}{n^2} \Var(y^T V_{\kappa} V_{\kappa}^T y) \\
-& = \frac{1}{n^2} \Var((\beta + \varepsilon)^T V_{\kappa} V_{\kappa}^T (\beta + \varepsilon)) \\
-& = \frac{1}{n^2} \Var(2 \beta^T V_{\kappa} V_{\kappa}^T \varepsilon + \varepsilon^T V_{\kappa} V_{\kappa}^T \varepsilon) \\
-& \leq \frac{1}{n^2}(4 \beta^T V_{\kappa} V_{\kappa}^T \beta + 2\kappa)
-\end{align}
-where the last inequality follows from standard properties of the Gaussian distribution.
-We now move on to showing the desired inequalities \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.
-For simplicity, we introduce the notation
-\begin{equation*}
-t(b) = b\sqrt{\frac{2\kappa}{n^2}}, ~~ \Delta = \norm{\beta}_2^2 - \frac{S_2(\beta;G)}{s_{\kappa}}.
-\end{equation*}
-
-\vspace{.2 in}
-
-\textit{Proof of~\eqref{eqn:graph_spectral_type_I_error}:} By Chebyshev's inequality,
-\begin{align*}
-\Pbb_{\beta = 0}\bigl(T_{\spec} \geq \frac{\kappa}{n} + t(b)\bigr)
-& \leq \Pbb_{\beta = 0}\bigl(\abs{T_{\spec} - \frac{\kappa}{n}} \geq t(b)\bigr) \\
-& \leq \frac{\Var_{\beta = 0}(T_{\spec})}{t(b)^2} = \frac{1}{b^2}.
-\end{align*}
-
-\vspace{.2 in}
-
-\textit{Proof of~\eqref{eqn:graph_spectral_type_II_error}:} Assumption~\eqref{eqn:fixed_graph_testing_critical_radius} implies
-\begin{equation*}
-\Delta \geq \frac{\norm{\beta}_2^2}{n} - \frac{C_n^2}{nC^2} \geq 2 t(b).
-\end{equation*}
-Then another application of Chebyshev's inequality gives us
-\begin{align*}
-\Pbb_{\beta}\bigl(T_{\spec} \leq \frac{\kappa}{n} + t(b)\bigr) & = \Pbb_{\beta}\bigl(T_{\spec} - \Ebb_{\beta}(T_{\spec}) \leq t(b) - \Delta \bigr) \\
-& \leq \Pbb_{\beta}\bigl(\abs{T_{\spec} - \Ebb_{\beta}(T_{\spec})} \leq \Delta - t(b) \bigr) \tag{since $\Delta \geq t(b)$}	\\
-& \leq \frac{\Var_{\beta}(T_{\spec})}{(\Delta - t(b))^2} \\
-& \leq 4\frac{\Var_{\beta}(T_{\spec})}{\Delta^2} \tag{since $\Delta \geq 2t(b)$} \\
-& \leq 4\frac{2\kappa/n^2 + 4\beta^T V_{\kappa} V_{\kappa}^T \beta /n}{\Delta^2}.
-\end{align*}
-We handle each summand in the numerator separately. For the first term, since $\Delta \geq 2 t(b)$, we have
-\begin{equation}
-\label{eqn:spectral_type_II_error_pf1}
-\frac{2\kappa}{n^2\Delta^2} \leq \frac{1}{2b^2}.
-\end{equation}
-
-For the second term, noting that $\Delta = \beta^T V_{\kappa} V_{\kappa}^T \beta$, we have
-\begin{align}
-\frac{\beta^T V_{\kappa} V_{\kappa}^T \beta/n}{\Delta^2} & \leq \frac{1}{n\Delta} \nonumber \\
-& \leq \frac{1}{2nt(b)} \nonumber \\
-& = \frac{1}{2b\sqrt{2\kappa}}, \label{eqn:spectral_type_II_error_pf2}
-\end{align}
-and combining~\eqref{eqn:spectral_type_II_error_pf1} and~\eqref{eqn:spectral_type_II_error_pf2} yields~\eqref{eqn:graph_spectral_type_II_error}.
-
-\subsubsection{Step 2: Bounding neighborhood graph functionals}
-
-To make use of Lemma~\ref{lem:fixed_graph_testing} we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$.
-In particular, we will show that for some constants $c_1,c_2,c_3$, the following statements:
-\begin{enumerate}[label=(E\arabic*)]
-	\item 
-	\label{event:discrete_sobolev_norm}
-	\textbf{Graph Sobolev norm:} When $\beta = (f(x_1),\ldots,f(x_n))$ for $f \in W^{1,2}(\mathcal{X};R)$:
-	\begin{equation}
-	\label{eqn:continuous_to_discrete_sobolev_norm}
-	S_2(\beta;G_{n,r}) \leq c_1 \cdot n^2 r^{d + 2} 
-	\end{equation}
-	\item 
-	\label{event:eigenvalue_tail_decay}
-	\textbf{Eigenvalue tail bound:} If $r = (\log^{a}n)\cdot(\log n/n)^{1/d}$ for any $a > 0$, then for all $\kappa \in [n]$:
-	\begin{equation}
-	\label{eqn:eigenvalue_tail_bound}
-	s_{\kappa} \geq c_2 \cdot \frac{\kappa^{2/d}}{n^{2/d}}
-	\end{equation}
-	\item 
-	\label{event:l2_norm}
-	\textbf{Empirical norm of $f$:} When $f \in W^{1,2}(\mathcal{X};R)$ and one of (a) $d = 1,2$, or (b) $d = 3$ and $\norm{f}_{\Leb^2}^2 \geq n^{-2/(d - 2)}$, then
-	\begin{equation}
-	\label{eqn:l2_to_empirical_norm}
-	\frac{1}{n}\sum_{i = 1}^{n}\beta_i^2 = \norm{f}_n^2 \geq c_3 \cdot \norm{f}_{\Leb^2}^2
-	\end{equation}
-\end{enumerate}
-
-each hold with probability $1 - o(1)$ as $n \to \infty$.
-
-\paragraph{Proof of~\eqref{eqn:continuous_to_discrete_sobolev_norm}:}
-
-Observe that
-\begin{align*}
-\frac{1}{n^2}\Ebb(\beta^T L \beta) = \Ebb\left(\frac{1}{n^2} \sum_{i,j = 1}^{n} (\beta_i - \beta_j)^2 A_{ij}\right) & = \frac{(n - 1)}{n} \int_{\Xset} \int_{\Xset} (f(x) - f(y))^2K_r(x,y) \,dP(y) \,dP(x) \\
-& \leq p_{\max}^2 \frac{(n - 1)}{n} \int_{\Xset} \int_{\X} (f(x) - f(y))^2K_r(x,y) \,dy \,dx.
-\end{align*}
-We will show that for any $f \in \mathcal{W}^{1,2}(\X;R)$
-\begin{equation}
-\label{eqn:nonlocal_continuous_sobolev_norm}
-\int_{\Xset} \int_{\Xset} (f(x) - f(y))^2K_r(x,y) \,dy \,dx \leq c r^{d + 2} \norm{f}_{W^{1,2}(\mathcal{X})}
-\end{equation}
-whence the desired result~\eqref{eqn:continuous_to_discrete_sobolev_norm} follows by Markov's inequality and the assumption $f \in W^{1,2}(\mathcal{X};R)$.
-
-We begin by dealing with complications due to smoothing at the boundary of $\mathcal{X} = [0,1]^d$.
-Let $U$ be any bounded open set such that $\Xset \subset U$ is compactly contained in $U$.
-Note that as $\partial \Xset$ is $C^1$, by Theorem \ref{thm:evans_extension} there exists $g \in W^{1,2}(\Reals^d)$ such that
-\begin{enumerate}[(i)]
-	\item
-	\label{eqn:evans_extension_1}
-	$g = f$, $P$-almost-everywhere in $\Xset$
-	\item 
-	$g$ has support within $V$, and  
-	\item 
-	\label{eqn:sobolev_1_bound_2}
-	$\norm{g}_{W^{1,2}(\Rd)} \leq c \norm{f}_{W^{1,2}(\Xset)}$ for a constant $c$ which depends only on $\Xset$.
-\end{enumerate}
-As a result of \ref{eqn:evans_extension_1},
-\begin{equation}
-\label{eqn:sobolev_1_bound_1}
-\int_{\Xset} \int_{\Xset} (f(x) - f(y))^2K_r(x,y) \,dy \,dx \leq \int_{\Rd} \int_{\Rd} (g(x) - g(y))^2K_r(x,y) \,dy \,dx.
-\end{equation}
-
-Next we smooth $g$, so that we may work with ordinary partial derivatives.
-We let $\eta \in C^{\infty}(\Rd)$ be given by
-\begin{equation*}
-\eta(x) :=
-\begin{cases}
-C \exp \left\{\frac{1}{\norm{x}^2 - 1}\right\}& \quad \textrm{if $\norm{x}_2 \leq 1$} \\
-0 & \quad \textrm{if $\norm{x}_2 \geq 1$}
-\end{cases}
-\end{equation*}
-where the normalizing constant $C > 0$ is chosen so that $\int_{\Rd} \eta dx = 1$. Let $\eta_r(x) := (1/r^d) \eta(x/r)$. Then, the mollification of $g$ by $\eta_r$ is given by
-\begin{align*}
-g^r & := g \ast \eta_r \\
-& = \int_{\Rd} \eta_r(x - y) g(y) dy
-\end{align*}
-(Refer to \textcolor{red}{Evans}, Appendix C, Theorem 7 for a proof that $g^r \in C^{\infty}(\Rd)$.)
-Adding and subtracting within \eqref{eqn:sobolev_1_bound_1}, we have
-\begin{align}
-\int_{\Rd} & \int_{\Rd} (g(x) - g(y))^2K_r(x,y) \,dy \,dx \\
-& \leq 3 \int_{\Rd} \int_{\Rd} \bigl((g(x) - g^r(x))^2 + (g^r(x) - g^r(y))^2 + (g^r(y) - g(y))^2\bigr)K_r(x,y) \,dy \,dx \nonumber \\
-& = 6 \int_{\Rd} \int_{\Rd} \bigl((g(y) - g^r(y))^2 K_r(x,y) \,dy \,dx + 2 \int_{\Rd} \int_{\Rd} \bigl((g^r(x) - g^r(y))^2 K_r(x,y) \,dy \,dx \label{eqn:sobolev_1_bound_5}
-\end{align}
-We deal with each summand individually, beginning with the first one. Let $\nu_d = \pi^{d/2}/\Gamma(d/2 + 1)$ denote the volume of a ball with unit radius in $\Rd$. Then, we have
-\begin{align}
-\int_{\Rd} \int_{\Rd} \bigl((g(y) - g^r(y))^2 K_r(x,y) \,dy \,dx & = \int_{\Rd} \int_{B(x,r)} \bigl((g(y) - g^r(y))^2\,dy \,dx \nonumber \\
-& = \int_{\Rd} \int_{B(y,r)} \bigl((g(y) - g^r(y))^2\,dx \,dy \tag{by Tonelli's Theorem} \nonumber \\
-& = \nu_d r^d \int_{\Rd}\bigl((g(y) - g^r(y))^2 \,dy \\
-& \leq \nu_d r^{d + 2} \int_{\Rd} \norm{\nabla g(y)}^2 \,dy \label{eqn:sobolev_1_bound_4}
-\end{align}
-where the last line follows from Lemma \ref{lem:poincare_mollify}.
-
-We now turn our attention to the second summand. Note that as $g^r \in C^{\infty}(\Rd)$, we may apply Theorem \ref{thm:taylor_expansion} and obtain
-\begin{align*}
-(g^r(x) - g^r(y))^2 & = \left(\int_{0}^{1} \nabla g^r(x + t(y - x)) \cdot (y - x) \,dt\right)^2 \\
-& \leq \int_{0}^{1} \bigl(\nabla g^r(x + t(y - x)) \cdot (y - x)\bigr)^2 \,dt  \tag{Jensen's inequality} \\
-& \leq \norm{y - x}^2 \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 \,dt. \tag{Cauchy-Schwarz inequality}
-\end{align*}
-As a result, we have
-\begin{align*}
-\int_{\Rd} \int_{\Rd} \bigl((g^r(x) - g^r(y))^2 K_r(x,y) \,dy) \,dx & \leq \int_{\Rd} \int_{\Rd} \norm{y - x}^2  \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 K_r(x,y) \,dt \,dy \,dx \\
-& \leq r^{2} \int_{\Rd} \int_{\Rd} \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 K_r(x,y) \,dt \,dy \,dx \\
-& = r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + t(y - x))}^2 K_r(x,y) \,dx \,dt \,dy \\
-& = r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + tz)}^2 K_r(z) \,dx \,dt \,dz \tag{$z = y - x$}
-\end{align*}
-
-where we write $K_r(z) = \mathbf{1}(\norm{z} \leq r)$ in an abuse of notation. Next, we note that
-\begin{equation*}
-\int_{\Rd} \norm{\nabla g^r(x + tz)}^2 \,dx = \int_{\Rd} \norm{\nabla g^r(x)}^2 \,dx \leq \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
-\end{equation*}
-with the inequality following from Lemma \ref{lem:gradient_mollify_commute}. Therefore,
-\begin{align}
-r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + tz)}^2 K_r(z) \,dx \,dt \,dz & \leq r^{2} \int_{\Rd} K_r(z) \int_{0}^{1} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx  \,dt \,dz \nonumber \\
-& = \nu_d r^{2 + d} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx. \label{eqn:sobolev_1_bound_3}
-\end{align}
-By \eqref{eqn:sobolev_1_bound_5}, \eqref{eqn:sobolev_1_bound_4} and \eqref{eqn:sobolev_1_bound_3}, we have that 
-\begin{equation*}
-\int_{\Rd} \int_{\Rd} (g(x) - g(y))^2 K_r(x,y) \,dx \,dy \leq 9 \nu_d r^{d + 2} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
-\end{equation*}
-Then by \ref{eqn:sobolev_1_bound_2}, $\int_{\Rd} \norm{\nabla g(x)}^2 \,dx \leq c \int_{\Xset} \norm{\nabla f(x)}^2 \,dx$, and the desired result of \eqref{eqn:sobolev_1_bound_1} follows.
-
-\paragraph{Proof of~\eqref{eqn:eigenvalue_tail_bound}:}
-
-For ease of notation we write $t = n^{1/d}$, and assume without loss of generality that $t$ is whole number. Let $\overline{G} = (\overline{V},\overline{E})$ be the lattice over $[0,1]^d$, formally
-\begin{align*}
-\overline{V} & = \bigl\{\overline{x}_k := \frac{1}{n}(k_1,\ldots,k_d): k \in [t]^d\bigr\} \\
-\overline{E} & = \bigl\{(\overline{x}_k, \overline{x}_{\ell}): \norm{k - \ell}_1 = 1\bigr\}.
-\end{align*}
-
-By Theorem 1.1 of \textcolor{red}{Garcia Trillos and Slepcev}, with probability at least $1 - n^{-1}$ there exists a bijection $\pi: \overline{V} \to X$ such that
-\begin{equation*}
-\max_{k \in [t]^d} \abs{\overline{x}_k - \pi(\overline{x}_k)} \leq c \left(\frac{\log n}{n}\right)^{1/d}
-\end{equation*}
-Since our choice $r = \log^a(n) \cdot (\log n/n)^{1/d}$ implies $r = \omega(\left(\frac{\log n}{n}\right)^{1/d})$, for $n$ sufficiently large with probability $1 - n^{-1} = 1 - o(1)$ we have
-\begin{equation}
-\label{eqn:eigenvalue_tail_bound_pf1}
-(\overline{x}_k,\overline{x}_{\ell}) \in \overline{E} \Longrightarrow (\pi(\overline{x}_k),\pi(\overline{x}_{\ell})) \in E.
-\end{equation}
-Let $\overline{L}$ be the Laplacian of $\overline{G}$, with eigenvalues $\lambda_1 < \cdots \leq \lambda_n$. If \eqref{eqn:eigenvalue_tail_bound_pf1} holds, then for all $x \in \Reals^n$,
-\begin{equation*}
-x^T \overline{L} x \geq x^T L x \Longrightarrow \lambda_j \leq s_j~~ \textrm{for all}~ j \in [n].
-\end{equation*}
-The eigenvalues of the $d$-dimensional lattice are explicitly known.
-In particular when $j = h^d$ for some whole number $h$, we have $s_j \geq \lambda_j = 4 d \sin^2(\pi h/2t) \geq \pi^2 h^2/ (4 t^2)$, with a similar inequality (for possible different constants) holding when $j^{1/d}$ is not integral; this shows~\eqref{eqn:eigenvalue_tail_bound}.
-
-\paragraph{Proof of~\eqref{eqn:l2_to_empirical_norm}:}
-
-To prove~\eqref{eqn:l2_to_empirical_norm} we will upper bound
-\begin{equation*}
-\mathbb{E}\biggl[\Bigl(\frac{1}{n}\sum_{i = 1}^{n}f^2(x_i)\Bigr)^2\biggr],
-\end{equation*}
-and then apply the Paley-Zygmund inequality (Lemma~\ref{lem:paley_zygmund}). We have
-\begin{align*}
-\mathbb{E}\biggl[\Bigl(\frac{1}{n}\sum_{i = 1}^{n}f^2(x_i)\Bigr)^2\biggr] & = \frac{(n-1)}{n}\mathbb{E}(f^2(x_1)) + \frac{\mathbb{E}[f^4(x_1)]}{n} \\
-& \leq p_{\max}^2 \left(\norm{f}_{\Leb^2}^2 + \frac{\norm{f}_{\Leb^4}^4}{n}\right) \\
-& \overset{(i)}{\leq} p_{\max}^2 \left(\norm{f}_{\Leb^2}^2 + c\frac{\norm{f}_{\Leb^2}^{(4 - d)}}{n}\right) \\
-& \overset{(ii)}{\leq}  p_{\max}^2 \left(\norm{f}_{\Leb^2}^2 + c\norm{f}_{\Leb^2}^2\right).
-\end{align*}
-Here $(i)$ follows from two facts, the first being the embedding $W^{1,2}(\mathcal{X};R) \subseteq \Leb^q(\mathcal{X};R')$ for $q = 2d/(d - 2)$ when $d > 2$ and $q = \infty$ when $d \leq 2$ and some constant $R'$, and the second being a standard result on interpolation of $\Leb^p$ spaces, namely that since $f \in \Leb^2(\mathcal{X}) \cap \Leb^q(\mathcal{X})$, then
-\begin{equation*}
-\norm{f}_{\Leb^4} \leq \norm{f}_{\Leb^2}^{\lambda} \cdot \norm{f}_{\Leb^q}^{1 - \lambda},~~\textrm{where}~ \frac{1}{4} = \frac{\lambda}{2} + \frac{1 - \lambda}{q}.
-\end{equation*} 
-Then $(ii)$ follows from our assumptions on $\norm{f}_{\Leb^2}^2$: either 
-\begin{enumerate}[(a)]
-	\item $d \leq 2$ and $\norm{f}_{\Leb^2}^2 \geq 1/n$ so $\norm{f}_{\Leb^2}^{(4 - d)} \leq \max(R^{4 - d}/n, \norm{f}_{\Leb^2}^2) \leq c\norm{f}_{\Leb^2}^2$, or
-	\item $d = 3$ and $\norm{f}_{\Leb^2}^2 \geq n^{-2/(d - 2)}$.
-\end{enumerate} 
-Applying Lemma~\ref{lem:paley_zygmund} to  $\norm{f}_n^2$ yields~\eqref{eqn:l2_to_empirical_norm}.
-
-\subsubsection{Step 3: Conclusion}
-
-We note that for all possible values of $X \in \Xset^n$, under the null hypothesis $f = f_0 = 0$ and therefore $\beta = (f(x_1),\ldots,f(x_n)) = 0$ as well.
-Therefore by~\eqref{eqn:graph_spectral_type_I_error}, we have the following bound on Type I error:
-\begin{equation}
-\Ebb_{f_0}(\phi_{\mathrm{spec}}) = \mathbb{E}(\mathbb{E}_{\beta = 0}(\phi_{\spec}) | X) \leq \frac{1}{b^2}.
-\end{equation}
-
-Now, we bound Type II error under the assumption $f \in W^{1,2}(\mathcal{X};R)$ and 
-\begin{equation}
-\label{eqn:critical_radius_1}
-\norm{f}_{\Leb^2}^2 \geq \epsilon^2 = c \cdot b \cdot (\log n)^{h(\beta,d)} \cdot n^{-4/(4 + d)}.
-\end{equation}
-We verify that the conditions required to apply Step 2 are satisfied.
-In particular, we choose $r = (\log^a n) \cdot (\log n/n)^{1/d}$, and observe that~\eqref{eqn:critical_radius_1} implies $\norm{f}_{\Leb^2}^2 \geq n^{-1}$ when $d \leq 2$ and $\norm{f}_{\Leb^2}^2 \geq n^{-2/(d - 2)}$ when $2 < d < 4$.
-We may therefore apply our conclusions in Step 2; namely, that for every possible choice of $f$ there exists a good set $\mathcal{E}_f \subseteq \Xset^n$ with $\Pbb(\mathcal{E}_f) \geq 1 - o(1)$ such that each of \eqref{eqn:continuous_to_discrete_sobolev_norm}, \eqref{eqn:eigenvalue_tail_bound}, and \eqref{eqn:l2_to_empirical_norm} hold for all $X \subseteq \mathcal{E}_f$.
-Choosing $\kappa = n^{2d/(4 + d)}$ to balance the squared bias and variance terms on the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius}, we have that for all $X \subseteq \mathcal{E}_f$
-\begin{align*}
-2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_2(\beta;G)}{ns_{\kappa}} & \leq 2bn^{-4/(4+d)} + \frac{n r^{d + 2}}{s_{\kappa}} \tag{by \eqref{eqn:continuous_to_discrete_sobolev_norm}} \\
-& \leq 2bn^{-4/(4+d)} + c\cdot\frac{n r^{d + 2}}{n^{4/(4+d)}} \tag{by \eqref{eqn:eigenvalue_tail_bound}} \\
-& \leq 2bn^{-4/(4+d)} + c\cdot(\log^{a(d + 2) + (1 + 2/d)}n)\cdot n^{-4/(4+d)} \\
-& \leq \frac{1}{10}\norm{f}_{\Leb^2} \tag{for a suitably large choice of $c$ in \eqref{eqn:critical_radius_1}}\\
-& \leq \frac{1}{n}\sum_{i = 1} \beta_i^2. \tag{by \eqref{eqn:l2_to_empirical_norm}}
-\end{align*}
-We conclude that for all $X \subseteq \mathcal{E}_f$, the inequality \eqref{eqn:fixed_graph_testing_critical_radius} is satisfied with respect to $\beta = (f(x_1),\ldots,f(x_n))$ and $G = G_{n,r}$.
-As a result the worst-case Type II error is bounded
-\begin{equation*}
-\sup_{\substack{f \in W^{1,2}(\mathcal{X;R}) \\ \norm{f}_{\Leb^2} \geq \epsilon}}\mathbb{E}_{f}(1 - \phi_{\spec}) \leq \sup_{\substack{f \in W^{1,2}(\mathcal{X;R}) \\ \norm{f}_{\Leb^2} \geq \epsilon}} \mathbb{E}\bigl[\mathbb{E}_{\beta}(1 - \phi_{\spec}|X \in \mathcal{E}_f)\bigr] + o(1) \leq \frac{3}{b} + o(1).
-\end{equation*}
-
-\subsection{Proof of Proposition~\ref{prop:L4_testing_rate}}
-
-Let $T_{\mathrm{mean}} = \frac{1}{n}\sum_{i = 1}^{n} y_i^2$. The expectation of $T_{\mathrm{mean}}$ is
-\begin{equation*}
-\Ebb(T_{\mathrm{mean}}) = \mathbb{E}(f^2(x_1)) + 1,
-\end{equation*}
-and the variance can be upper bounded
-\begin{equation*}
-\Var(T_{\mathrm{mean}}) \leq \frac{1}{n}(3 + p_{\max} \norm{f}_{\Leb^4}^4 + p_{\max}\norm{f}_{\Leb^2}^2) = \frac{c}{n}.
-\end{equation*}
-Since $\mathbb{E}(f^2(x_1)) \geq p_{\min} \epsilon^2$, when $\epsilon^2 \gtrsim b\cdot n^{-1/2}$ we can apply Chebyshev's inequality to obtain the claimed result.
-\section{Auxiliary Results}
-
-\begin{theorem}[\textcolor{red}{Evans} Chapter 5.4, Theorem 1]
-	\label{thm:evans_extension}
-	Assume $U$ is bounded and $\partial U$ is $C^1$.
-Select a bounded open set $V$ such that $U \subset \subset V$ ($U$ is compactly contained in $V$).
-Then there exists a bounded linear operator $E: W^{1,2}(U) \to W^{1,2}(\Rd)$ such that for each $u \in W^{1,2}(U)$:
-	\begin{enumerate}
-		\item $Eu = u$ a.e. in $U$,
-		\item $Eu$ has support within $V$, and 
-		\item 
-		\begin{equation*}
-		\norm{Eu}_{W^{1,2}(\Rd)} \leq C \norm{u}_{W^{1,2}(\Rd)}
-		\end{equation*}
-		the constant $C$ depending only on $U$ and $V$.
-	\end{enumerate}
-\end{theorem}
-
-For $u \in W^{1,2}(\Rd)$ and $x \in \Rd$, write $\nabla u(x) = (D^{e_1}(x),\ldots,D^{e_d}(x)$, for the gradient of $u$.
-
-\begin{theorem}[\textcolor{red}{Evans} Chapter 5.8.1, Theorem 2]
-	\label{thm:evans_poincare}
-	There exists a constant $C$, depending only on $d$, such that
-	\begin{equation*}
-	\norm{u - u^r}_{L^2(B(x,r))} \leq Cr\norm{\nabla u}_{L^2(B(x,r))}
-	\end{equation*}
-\end{theorem}
-
-\begin{theorem}[Taylor expansion.]
-	\label{thm:taylor_expansion}
-	For any function $u \in C^1$, and any $x,y \in \Rd$, 
-	\begin{equation*}
-	u(y) - u(x) = \int_{0}^{1} \nabla(u(x + t(y - x))) \cdot (y - x) \,dt
-	\end{equation*}
-\end{theorem}
-
-\begin{lemma}
-	\label{lem:poincare_mollify}
-	For any function $g \in W^{1,2}(\Rd)$ compactly supported in a bounded open set $V \subset \Rd$, we have
-	\begin{equation}
-	\label{eqn:poincare_mollify}
-	\int_{\Rd} (g(x) - g^r(x))^2 \,dx \leq r^2 \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
-	\end{equation}
-\end{lemma}
-\begin{proof}
-	This Lemma is essentially a reproduction of part of the proof of the Rellich-Kondrachov Compactness Theorem from \textcolor{red}{Evans}.
-Note that it is sufficient to prove in the case when $g$ is smooth.
-To see this, for the moment assume \eqref{eqn:poincare_mollify} holds for all $u \in C^{\infty}(V)$, and let $g \in W^{1,2}(V)$.
-By Theorem \ref{thm:local_approx_smooth_functions}, we may take a sequence $g_m \in C^{\infty}(V)$ such that
-	\begin{equation*}
-	\norm{g - g_m}_{L^2(V)} \to 0, \quad \textrm{and} \quad \norm{\nabla g - \nabla g_m}_{L^2(V)} \to 0,
-	\end{equation*}
-	Then we have
-	\begin{align*}
-	\int_{\Rd} (g(x) - g^r(x))^2 \,dx & \leq \int_{\Rd} (g(x) - g_m(x))^2 \,dx + \int_{\Rd}(g_m(x) - g_m^r(x))^2 \,dx + \int_{\Rd}(g_m^r(x) - g^r(x))^2 \,dx \\
-	& \leq \int_{\Rd} (g(x) - g_m(x))^2 \,dx + r^2 \int_{\Rd} \norm{\nabla g_m(x)}^2 \,dx + \int_{\Rd}(g_m^r(x) - g^r(x))^2 \,dx
-	\end{align*}
-	and taking the limit as $m$ goes to infinity, the right hand side converges to $\int_{\Rd} \norm{\nabla g(x)}^2 \,dx$.
-
-	It remains to show \eqref{eqn:poincare_mollify} in the case where $g$ is smooth. In this case, we have
-	\begin{align*}
-	g^{r}(x) -  g(x) & = \frac{1}{r^d} \int_{B(x,r)} \kappa \left(\frac{x - z}{r}\right)\bigl(g(z) - g(x)\bigr) \,dz \\
-	& = \int_{B(0,1)} \kappa(y) \bigl(g(x - ry) - g(x)\bigr) \,dy \\
-	& = \int_{B(0,1)} \kappa(y) \int_{0}^{1} \frac{d}{dt} \bigl(g(x - try) \bigr) \,dt \,dy \\
-	& = -r \int_{B(0,1)} \kappa(y) \int_{0}^{1} \bigl(\nabla g (x - try) \bigr) \cdot y \,dt \,dy.
-	\end{align*} 
-	Therefore, by Jensen's and Cauchy-Schwarz inequalities, we have
-	\begin{align*}
-	\int_{\Rd} (g(x) - g^r(x))^2 \,dx & \leq r^2 \int_{\Rd} \int_{B(0,1)} \kappa(y) \int_{0}^{1} \norm{\nabla g (x - try)}^2 \norm{y}^2 \,dt \,dy \,dx \\
-	& \leq r^2 \int_{B(0,1)} \kappa(y) \int_{0}^{1} \int_{\Rd} \norm{\nabla g (x - try)}^2 \,dx \,dt \,dy \\
-	& = r^2 \int_{\Rd} \norm{\nabla g(z)}^2 \,dz
-	\end{align*}
-\end{proof}
-
-The following theorem is Theorem 1 in Section 5.3 of \textcolor{red}{Evans}.
-\begin{theorem}[Local approximation by smooth functions.]
-	\label{thm:local_approx_smooth_functions}
-	Assume $U$ is bounded, and $u \in W^{k,p}(U)$ for some $1 \leq p < \infty$. Then there exists functions $u_m \in C^{\infty}(U) \cap W^{k,p}(U)$ such that
-	\begin{equation*}
-	\norm{u_m - u}_{W^{k,p}(U)} \overset{m}{\to} 0.
-	\end{equation*}
-\end{theorem}
-
-\begin{lemma}
-	\label{lem:gradient_mollify_commute}
-	For any $u \in W^{1,2}(\Rd)$, we have
-	\begin{equation*}
-	\int_{\Reals^d} \norm{\nabla u^r(x)}_2^2 \,dx \leq \int_{\Reals^d} \norm{\nabla u(x)}_2^2 \,dx
-	\end{equation*}
-\end{lemma}
-\begin{proof}
-	Observe that $\norm{\nabla u(x)}_2^2 = \sum_{j = 1}^{d} (D^{e_j}u(x))^2 $. Therefore,
-	\begin{align*}
-	\int_{\Rd} \norm{\nabla u^r(x)}_2^2 \,dP(x) & = \sum_{j = 1}^{d} \int_{\Rd} (D^{e_j}u^r(x))^2 \,dx \\
-	& = \sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u)^r(x))^2 \,dx
-	\end{align*}
-	where the second equality follows from equation (1) in Section 5.3 of \textcolor{red}{Evans}. Then, for any $v \in L^2(\Rd)$, we have that
-	\begin{align*}
-	\abs{v^r(x)} & = \int_{\Rd} \kappa_r^{1/2}(x - y) \kappa_r^{1/2}(x - y) v(y) \,dy \\
-	& \leq \left(\int_{\Rd} \kappa_r^(x - y) \,dy\right)^{1/2} \left(\int_{\Rd} \kappa_r(x - y) v^2(y) \right)^{1/2} \\
-	& = \left(\int_{\Rd} \kappa_r(x - y) v^2(y) \right)^{1/2}
-	\end{align*}
-	and therefore
-	\begin{align*}
-	\int_{\Rd} (v^r(x))^2 \,dx & \leq \int_{\Rd} \int_{\Rd } \kappa_r(x - y) v^2(y) \,dy \,dx \\
-	& = \int_{\Rd} \int_{\Rd } \kappa_r(x - y) v^2(y) \,dy \,dx \\
-	& = \int_{\Rd} v^2(y) \,dy
-	\end{align*}
-
-	Applying this to $D^{e_j}u \in L^2(\Rd)$, we have that
-	\begin{equation*}
-	\sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u)^r(x))^2 \,dx \leq \sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u(x))^2 \,dx = \int_{\Rd} \norm{\nabla u(x)}^2 \,dx.
-	\end{equation*}
-\end{proof}
-
-\begin{lemma}
-	\label{lem:paley_zygmund}
-	Let $0 < p < q$, and let $Z$ be a non-negative random variable such that $\mathbb{E}(Z^q) < \infty$. Then, for all $0 \leq \lambda \leq 1$,
-	\begin{equation}
-	\label{eqn:paley_zygmund}
-	\mathbb{P}(Z > \lambda \mathbb{E}(Z^p)) \geq \left[(1 - \lambda^p) \frac{\mathbb{E}(Z^p)}{(\mathbb{E}(Z^q))^{p/q}}\right]^{\frac{q}{q - p}}
-	\end{equation}
-\end{lemma}
-
+where the last inequality can be verified by a quick calculation in light of $\kappa = n^{2d/(4s + d)}$ and $r \leq n^{-4/((2+d)(4s + d))}$.
+Since we've already shown $\lambda_{\kappa}(G_{n,r}) \geq \lambda_{\kappa}(H)$ with probability $1 - o(n^{-1})$, this completes the proof of Lemma~\ref{lem:higher_order_eigenvalue_tail_bound}.
 \end{document}

### 2020-01-09 notes/weekly_notes/12.20.19.NotesForWeek.tex
added weekly notes for break

@@
 \begin{document}
-\title{Notes for Week 12/7/19 - 12/20/19}
+\title{Notes for Week 12/7/19 - 12/29/19}
 \author{Alden Green}
@@
 \begin{equation*}
-\int z^j K(z) = 0,~~\textrm{$j = 1,2$.}
+\int z^j K(z) = 0~~\textrm{for $j = 1,2$.}
 \end{equation*}
@@
 	\label{lem:3rd_order_roughness_functional}
-	Suppose $f \in C^{3}(L)$ for some $L > 0$, and further suppose $p \in C^{2}(L)$. For any kernel $K$ which satisfies the previous conditions, and any $r(n) \geq n^{-1/(3+d)}$, we have that
+	Suppose $f \in C^{3}(L)$ for some $L > 0$, and further suppose $p \in C^{2}(L)$. For any kernel $K$ which satisfies the previous conditions, and any $r(n) \geq n^{-1/(4+d)}$, we have that
 	\begin{equation*}
-	\Ebb(R_{s,n}(f)) = O(L^2)
+	\Ebb(R_{s,n}(f)) = O(1)
 	\end{equation*} 
@@
 	Suppose $f \in C^{s}(L)$ for some $s,L > 0$, and further suppose $p \in C^{\ell}(L)$ for $\ell = \floor{s}$ the largest integer strictly less than $s$.
-For any order-$\ell$ kernel $K$ and any $r(n) \geq n^{-1/(4+d)}$, we have that
+For any order-$\ell$ kernel $K$ and any $1 \geq r(n) \geq n^{-1/(2(s - 1)+d)}$, we have that
 	\begin{equation*}
-	\Ebb(R_{s,n}(f)) = O(r^{2s})
+	\Ebb(R_{s,n}(f)) = O(1)
 	\end{equation*} 
@@
 When $s \geq 3$ the graph Laplacian approximates higher-order derivatives by composing difference operators.
-The crudeness of this approximation means that in Lemma~\ref{lem:roughness_functional_expectation} we allow for a slower scaling of $r(n)$ than we did Lemma~\ref{lem:2nd_order_roughness_functional}.
+The crudeness of this approximation means that in Lemma~\ref{lem:roughness_functional_expectation} we allow for a slower scaling of $r(n)$ than we did in Lemmas~\ref{lem:2nd_order_roughness_functional} and~\ref{lem:3rd_order_roughness_functional}.
 At a high level, however, we prove Lemma~\ref{lem:roughness_functional_expectation} in a similar fashion to Lemma~\ref{lem:2nd_order_roughness_functional}.
 We first break the sum into cases according to the number of unique indices, then bound the expectation case by case.
-To simplify our statements, for index vectors $k,\ell \in [n]^q$ we will use the following notation:
-\begin{equation*}
-\textrm{Unq}(k \setminus \ell) = \set{x \in k: x \not\in \ell, \textrm{$x$ appears in $k$ exactly once,}}~~ k \bigcup \ell = \set{x: x \in k ~\textrm{or}~ x \in \ell}
-\end{equation*}
-Thus $\textrm{Unq}(k \setminus \ell)$ measures the number of indices in $k$ which appear exactly once in $k$ and never in $\ell$, and $k \bigcup \ell$ measures the total number of distinct indices.
-
-The following Lemma will be the workhorse which supplies a sufficient bound in all cases.
+The following Lemma will be the workhorse which supplies a sufficient bound in each case.
 \begin{lemma}
@@
 	\begin{equation}
-	\Ebb(D_kf(x_i) D_kf(x_i)) =
+	\label{eqn:expected_difference_operators_1}
+	\Ebb(D_kf(x_i) D_\ell f(x_i)) =
 	\begin{cases*}
-	O(r^{2} r^{d(\abs{k \cup \ell} - 2q)}), & ~~\textrm{if $\Unq(k\setminus\ell) = \emptyset$ and $\Unq(\ell\setminus k) = \emptyset$}~ 
-	\\
-	O(r^{2q + 1} r^{\abs{k \cup \ell}} r^{d(\abs{k \cup \ell} - 2q)}), & ~~\textrm{if $\Unq(k\setminus\ell) \neq \emptyset$ or $\Unq(\ell\setminus k) \neq \emptyset$} \\
-	O(r^{4q} r^{d(\abs{k \cup \ell} - 2q)}), & ~~\textrm{if $\Unq(k\setminus\ell) \neq \emptyset$ and $\Unq(\ell\setminus k) \neq \emptyset$}
+	O(r^{2s}), & ~~\textrm{if all indices are distinct} \\
+	O(r^{2} r^{d(\abs{k \cup \ell \cup i} - (2q + 1))}), & ~~\textrm{otherwise}~ 
+	\end{cases*}
+	\end{equation}
+	Additionally, we have
+	\begin{equation}
+	\label{eqn:expected_difference_operators_2}
+	\Ebb(d_iD_kf(x_j) d_iD_\ell f(x_j)) =
+	\begin{cases*}
+	O(r^{2s}), & ~~\textrm{if all indices are distinct} \\
+	O(r^{2} r^{d(\abs{k \cup \ell \cup i \cup j} - (2q + 2))}), & ~~\textrm{otherwise}~ 
 	\end{cases*}
@@
 \section{Proofs}
+
+\subsection{Proof of Lemma~\ref{lem:roughness_functional_expectation}}
+We will take $s$ to be even, as the proof when $s$ is odd follows essentially the same steps. Now when $s$ is even, we have the decomposition
+\begin{equation*}
+R_{s,n}(f) = \frac{1}{n^{s+1}r^{2s}} \sum_{i = 1}^{n} \sum_{k \in [n]^q} \sum_{\ell \in [n]^q} D_kf(x_i) D_{\ell}f(x_i)
+\end{equation*}
+For given index vectors $k,\ell$ and index $i$, let $I = \abs{k \cup \ell \cup i}$ be the total number of unique indices. We separate our analysis into cases based on the magnitude of $I$.
+
+When $I < s + 1$, there is at least one repeated index, and by Lemma~\ref{lem:expected_difference_operators} we have that
+\begin{equation*}
+\Ebb(D_kf(x_i)D_{\ell}f(x_i)) = O(r^2r^{d(I - (s + 1))}).
+\end{equation*}
+There are $O(n^{I})$ terms in $R_{s,n}(f)$ with exactly $I$ distinct indices.
+Therefore, the total contribution of such terms to the sum is $O(r^{-2(s - 1)}r^{d(I - (s + 1))}n^{I - (s + 1)}r)$.
+Since $r(n) \geq n^{-1/d}$, this increases with $I$.
+Taking $I = s$ to be as large as possible such that there is at least one repeated index, the contribution of these terms to the sum is $O(r^{-(2(s - 1) + d)}n^{-1}) = O(1)$, with the equality following by the restriction $r \geq n^{-1/(2(s - 1) + d)}$.
+
+When $I = s + 1$, every index in $\ell,k$ and $i$ is unique. Therefore by Lemma~\ref{lem:roughness_functional_expectation},  
+\begin{equation*}
+\Ebb(D_kf(x_i)D_{\ell}f(x_i)) = O(r^{2s}).
+\end{equation*}
+Since there are $O(n^{s+1})$ such terms, we have that the total contribution of these terms is $O(n^{s + 1}r^{2s}n^{-(s + 1)}r^{-2s}) = O(1)$.
+
 \subsection{Proof of Lemma~\ref{lem:2nd_order_roughness_functional}}
@@
 \section{Proof of Lemma~\ref{lem:3rd_order_roughness_functional}}
+We prove Lemma~\ref{lem:3rd_order_roughness_functional} when $d = 1$ and $r(n) \geq n^{-1/5}$.
 
-\textcolor{red}{TODO}: Copy proof of Lemma~\ref{lem:3rd_order_roughness_functional} from handwritten notes.
+\textcolor{red}{TODO:} Prove for all dimensions $d$, by replacing derivatives with partial derivatives.
 
+From~\eqref{eqn:roughness_functional_representation_odd}, we have
+\begin{align*}
+R_{3,n}(f) & = \frac{1}{2n^2r^2}\sum_{i,j = 1}^{n}\left(\frac{1}{(nr^2)}\sum_{k \in n}\bigl(D_kf(x_i) - D_kf(x_j)\bigr)\right)^2K_r(x_i,x_j) \\
+& = \frac{1}{2n^4r^6}\sum_{i,j = 1}^{n} \sum_{k,\ell = 1}^{n} D_{i\ell}f(x_j) \cdot d_i(D_kf(x_j)).
+\end{align*}
+We separate our analysis into cases based on the distinct indices in the previous summation.
+
+\paragraph{Case 1: $i = k = \ell$.}
+In this case, we can rewrite the summand as
+\begin{align*}
+D_iif(x_j) d_iD_if(x_j) & = (d_if(x_j))^2 K_r(x_i,x_j)^3 \\
+& \leq 4L^2r^2 \abs{K_r(x_i,x_j)^3}
+\end{align*}
+Therefore,
+\begin{align*}
+\Ebb(D_iif(x_j) d_iD_if(x_j)) & \leq 4L^2 r^2 \Ebb\bigl(\abs{K_r(x_i,x_j)^3}\bigr)\\
+& \leq 4CL^3, \tag{Lemma~\ref{lem:expected_kernel}}
+\end{align*}
+and as a result
+\begin{equation*}
+\frac{1}{2n^4r^6}\sum_{i,j = 1}^{n} \Ebb(D_{ii}f(x_j) \cdot d_i(D_if(x_j))) \leq \frac{2CL^3}{n^2r^6} \leq \frac{2CL^3}{n^{4/5}}.
+\end{equation*}
+
+\paragraph{Case 2: $i = \ell \neq k$.}
+In this case, we can rewrite the summand as $-(D_if(x_j))(d_iD_kf(x_j))K_r(x_i,x_j)$. We have that
+\begin{equation*}
+\abs{D_if(x_j)} \leq 2LrK(x_i,x_j),~~\textrm{and}~ \abs{d_iD_kf(x_j)} \leq 2Lr \bigl(\abs{K_r(x_k,x_i)} + \abs{K_r(x_{\ell},x_i)}\bigr)
+\end{equation*}
+Therefore, we can upper bound
+\begin{equation*}
+\abs{D_if(x_j)d_iD_kf(x_j)K_r(x_i,x_j)} \leq 4Lr^2K(x_i,x_j)^2\bigl(\abs{K_r(x_k,x_i)} + \abs{K_r(x_{\ell},x_i)}\bigr)
+\end{equation*}
+and the resultant expectation satisfies
+\begin{equation*}
+\Ebb\bigl(\abs{D_if(x_j)d_iD_kf(x_j)K_r(x_i,x_j)}\bigr) \leq 8CL^4r.
+\end{equation*}
+Since there are order $n^3$ terms in the sum for which $i = k \neq \ell$, we have
+\begin{equation*}
+\frac{1}{2n^4r^4}\sum_{\ell\neq k = i \neq j\neq }^{n} \Ebb(D_{ii}f(x_j) \cdot d_i(D_{\ell}f(x_j))) \leq \frac{8 C L^3}{r^5n} \leq 8 C L^3.
+\end{equation*}
+Note that by symmetry, the same analysis applies to the case $i = \ell \neq k$.
+\paragraph{Case 3: $i \neq k = \ell$.}
+In this case, we can rewrite the summand as $(d_{i}D_kf(x_j))^2K_r(x_i,x_j)$. We bound the order of magnitude of the composed difference operators using Taylor expansion:
+\begin{align*}
+\abs{d_{i}D_kf(x_j)} & = \abs{(f(x_k) - f(x_i))K_r(x_k,x_i) - (f(x_k) - f(x_j))K_r(x_k,x_j)} \\
+& \leq \abs{(f(x_k) - f(x_i))}\abs{K_r(x_k,x_i)} + \abs{(f(x_k) - f(x_j))}\abs{K_r(x_k,x_ij)} \\
+& \leq 2L\left(\abs{x_k - x_j} \abs{K_r(x_k,x_i)} + \abs{x_k - x_i} \abs{K_r(x_k,x_i)}\right) \\
+& \leq 2Lr \bigl(\abs{K_r(x_k,x_i)} + \abs{K_r(x_{\ell},x_i)}\bigr)
+\end{align*}
+Taking expectation with respect to $x_k$, we have
+\begin{align*}
+\Ebb(\bigl(d_{i}D_kf(x_j)\bigr)^2|x_i,x_j) & \leq 8L^2r^2\left(\int K_r(x,x_i)^2 p(x)\,dx + \int K_r(x,x_j)^2 p(x)\,dx\right) \\
+& \leq 16L^3r^{2} \int \left(\frac{1}{r}K(z/r)\right)^2 \,dz \tag{$p \in C^2(L)$} \\
+& \leq 16L^3r \int \bigl(K(t)\bigr)^2 \,dt \tag{change of variables} \\
+& \leq 16CL^3r. \tag{$K$ compactly supported and bounded}
+\end{align*}
+As a result, 
+\begin{align*}
+\Ebb((d_{i}D_kf(x_j))^2K_r(x_i,x_j)) & = \Ebb(\Ebb((d_{i}D_kf(x_j))^2|x_i,x_j)K_r(x_i,x_j)) \\
+& \leq 16CL^3r\Ebb(\abs{K_r(x_i,x_j)}) \\
+& \leq 16CL^3r.
+\end{align*}
+Since there are not quite $n^3$ terms in the sum with $i \neq k = \ell$, and since $r \geq n^{-1/5}$, 
+\begin{equation*}
+\frac{1}{2n^4r^6}\sum_{i \neq j \neq k = \ell \neq i}^{n} \Ebb(D_{ik}f(x_j) \cdot d_i(D_kf(x_j))) \leq \frac{8 C L^3}{r^5n} \leq 8 C L^3.
+\end{equation*}
+
+\paragraph{Case 4: $i,k,\ell$ all distinct.}
+We rewrite the summand as $d_i(D_{\ell}f(x_j)) d_i(D_{k}f(x_j))K_r(x_i,x_j)$. Noting that
+\begin{equation*}
+d_i(D_{k}f(x_j)) = [f(x_k) - f(x_i)]K_r(x_k,x_i) -  [f(x_k) - f(x_j)]K_r(x_k,x_j)
+\end{equation*}
+by Lemma~\ref{lem:leading_term} we have that
+\begin{equation*}
+\abs{\Ebb(d_i(D_{k}f(x_j))|x_i,x_j)} \leq 2Lr^3.
+\end{equation*}
+Therefore,
+\begin{align*}
+\abs{\Ebb(d_i(D_{\ell}f(x_j)) d_i(D_{k}f(x_j))K_r(x_i,x_j))} & \leq \Ebb\biggl(\abs{\Ebb(d_k(D_{\ell}f(x_j))|x_i,x_j)\cdot\Ebb(d_k(D_{\ell}f(x_j))|x_i,x_j)\cdot K_r(x_k,x_j)}\biggr) \\
+& \leq 4L^2r^6 \Ebb(\abs{K_r(x_k,x_j)}) \leq 4CL^2r^6.
+\end{align*}
+Since there are order $n^4$ terms in the summation for which $i,j,k$ and $\ell$ are all distinct, we have that
+\begin{equation*}
+\frac{1}{2n^4r^6} \sum_{i \neq j \neq k \neq \ell}^{n} \Ebb(D_{ik}f(x_j) d_iD_{\ell}f(x_j) K_r(x_i,x_j)) \leq 4C.
+\end{equation*}
+The above four cases cover all terms in the summation, and so we have proved Lemma~\ref{lem:3rd_order_roughness_functional}.
 \section{Proof of Lemma~\ref{lem:expected_difference_operators}}
 
-\textcolor{red}{TODO}: Prove Lemma~\ref{lem:expected_difference_operators}.
+\subsection{Proof of~\eqref{eqn:expected_difference_operators_1}}
 
+Let $k,\ell \in [n]^q$ be index vectors. We first prove the desired bound in the case when some indices are repeated, and then the desired bound in the case when all indices are distinct.
+
+\paragraph{Repeated indices.}
+
+Since $f \in C^1(L)$, for all index vectors $k,\ell \in [n]^q$ and indices $i \in [n]$, the product of iterated difference operators $D_kf(x_i) D_{\ell}f(x_i)$ satisfies
+\begin{equation*}
+\abs{D_kf(x_i) D_{\ell}f(x_i)} \leq 4^{q} L^2 r^{2 - 2q}
+\end{equation*}
+Moreover $D_kf(x_i) D_{\ell}f(x_i)$ will equal zero if there exists $x_j, j \in k \cup \ell \cup i$ such that
+\begin{equation*}
+\norm{x_j - x_h} > r,~~\textrm{for all $h \in k \cup \ell \cup i$}
+\end{equation*}
+Therefore $D_kf(x_i) D_{\ell}f(x_i)$ is nonzero with probability $O(r^{\abs{k \cup \ell \cup i} - 1})$, which along with the boundedness of $D_kf(x_i) D_{\ell}f(x_i)$ implies the claimed result.
+
+\paragraph{All indices distinct.}
+
+By the law of iterated expectation, we have
+\begin{align}
+\Ebb\left[D_kf(x_i)D_{\ell}f(x_i)\right] & = \Ebb\left[\Ebb\left(D_kf(x)|x_i = x\right) \Ebb\left(D_{\ell}f(x)|x_i = x\right)\right] \nonumber \\
+& = \Ebb\left[\Ebb\left(D_kf(x)|x_i = x\right)^2\right] \label{eqn:expected_difference_operators_pf1}
+\end{align}
+Again applying the law of iterated expectation, letting $g$ be the function given by $g(x) = \Ebb(D_1f(x))$, we have
+\begin{align*}
+\Ebb(D_kf(x)) & = \Ebb(\Ebb(D_{k}f(x)|x_{k_1},\ldots,x_{k_{q-1}})) \\
+& = \Ebb((D_{k_1\cdots k_{q - 1}}g)(x)).
+\end{align*}
+By Lemma~\ref{lem:leading_term}, we have that $g(x) = O(r^s)$ for all values of $x$, and therefore
+\begin{equation*}
+(D_{k_1\cdots k_{q_1}}g)(x) = O(r^{s - d(q - 1)})
+\end{equation*}
+Since $(D_{k_1\cdots k_{q_1}}g)(x)$ is nonzero with probability $O(r^{d(q - 1)})$, we have that 
+\begin{equation*}
+\Ebb(D_kf(x)) = \Ebb((D_{k_1\cdots k_{q - 1}}g)(x)) = O(r^s),
+\end{equation*}
+and the desired result follows from~\eqref{eqn:expected_difference_operators_pf1}.
+
+\subsection{Proof of~\eqref{eqn:expected_difference_operators_2}}
+
+The proof of~\eqref{eqn:expected_difference_operators_2} is essentially the same, and we do not reproduce it here.
 \section{Auxiliary Results.}
@@
 
+\begin{lemma}
+	\label{lem:expected_kernel}
+	When $K$ is a bounded and compactly supported kernel, then for any $s$ and for all $x \in \Rd$
+	\begin{equation*}
+	\Ebb(\abs{K(x_i,x)}^s) \leq CLr^{(1 - s)d}
+	\end{equation*}
+\end{lemma}
+
+\begin{lemma}
+	\label{lem:leading_term}
+	When $f \in C^s(L)$ and $p \in C^{s - 1}(L)$, we have that 
+	\begin{equation*}
+	\Ebb(\left(f(x_{\ell}) - f(x)\right)K_r(x_{\ell},x)) \leq Lr^s.
+	\end{equation*}
+\end{lemma}
+
 \end{document}

### 2020-01-13 notes/graph_testing.tex
added minimax regression testing rate for higher order Holder classes

@@
 \begin{equation*}
-\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, \norm{f - f_0}_2 \geq \epsilon} \Ebb_f(1 - \phi).
+\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, \norm{f - f_0}_{\Leb^2} \geq \epsilon} \Ebb_f(1 - \phi).
 \end{equation*}
@@
 \end{theorem}
-The testing rate~\eqref{eqn:sobolev_testing_rate} matches the minimax critical radius up to a factor of $\log^{h(a,d)}n$.
 
@@
 Note that when $d < 4$ the rate~\eqref{eqn:sobolev_testing_rate} is sharper than \eqref{eqn:L4_testing_rate}.
+
+The minimax testing rate over the higher-order smoothness classes $W_d^{s,2}(\Xset;L)$ and $C_d^s(\Xset;L)$ is known to be $n^{-4s/(4s + d)}$ when $4s < d$.
+Hence the testing rate~\eqref{eqn:sobolev_testing_rate} matches the minimax critical radius when $s = 1$ up to a factor of $\log^{h(a,d)}n$.
+The eigenvector projection test is also minimax optimal over the balls in higher order Holder spaces $C_d^s(\Xset;L)$.
+
+\begin{theorem}
+	\label{thm:higher_order_holder_testing_rate}
+	Let $b \geq 1$ be a fixed constant, and let $d$ and $s$ be positive integers such that $d < 4s$.
+Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p \in C^{s-1}(\mathcal{X};R)$ bounded above and below by constants, i.e
+	\begin{equation*}
+	0 < p_{\min} < p(x) < p_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
+	\end{equation*}
+	Then the following statement holds: if the test $\phi_{\spec}$ is performed with parameter choices 
+	\begin{equation*}
+	\textrm{$K_r$ the uniform kernel with}~ n^{-1/(2(s - 1) + d)} \leq r(n) \leq n^{-4/((4s + d)(2+d))}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	\end{equation*}
+	then there exists constants $c_1,c_2$ which may depend on $d,R,$ and $s$ but are independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
+	\begin{equation}
+	\label{eqn:higher_order_holder_testing_rate}
+	\epsilon^2 \geq c_1 \cdot b^2 \cdot n^{-4s/(4s + d)}
+	\end{equation}
+	the worst-case risk is upper bounded
+	\begin{equation}
+	\label{eqn:higher_order_holder_testing_rate_1}
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; C_d^{s}(\mathcal{X};R)) \leq \frac{c_2}{b}.
+	\end{equation}
+\end{theorem}
 
@@
 
-\section{\textcolor{red}{In Progress}}
-
-We say $K_{r}$ is an order-$s$ kernel when 
-\begin{equation*}
-\int x^j K_{r}(x) \,dx = 0, \textrm{for $j = 1,\dots,s$}
-\end{equation*}
+\section{\textcolor{red}{In Progress Theorems}}
 
@@
 	\label{thm:higher_order_sobolev_testing_rate}
-	Let $b,s,d \geq 1$ and $a > 0$ be fixed constants such that $d < 4s$.
-Suppose that $\Pbb$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p$ bounded above and below by constants, i.e
+	Let $b,s$ and $d \geq 1$ be fixed constants such that $d < 4s$.
+Suppose that $\Pbb$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p \in C^{s-1}(\Xset;R)$ bounded above and below by constants, i.e
 	\begin{equation*}
@@
 	\begin{equation*}
-	\textrm{$K_r$ an order-$s$ kernel with}~ r = \log^a n \cdot \left(\frac{\log n}{n}\right)^{1/d}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
-	\end{equation*}
-	then there exists a constant $c$ which may depend on $d,R,p_{\max},b$ and $a$ but is independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
+	\textrm{$K_r$ the uniform kernel,}~ n^{-1/(2(s-1) + d)} \leq r(n) \leq n^{-4/((4s + d)(2+d))}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	\end{equation*}
+	then there exists constants $c_1,c_2$ which may depend on $d,R$ and $s$ but is independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
 	\begin{equation}
 	\label{eqn:higher_order_sobolev_testing_rate}
-	\epsilon^2 \geq c \cdot b \cdot n^{-4/(4s + d)} (\log n)^{h(a,d)}
+	\epsilon^2 \geq c \cdot b^2 \cdot n^{-4/(4s + d)}
 	\end{equation}
@@
 	\label{eqn:higher_order_sobolev_testing_rate_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; \mathcal{W}^{s,2}(\mathcal{X};R)) \lesssim \frac{1}{b}.
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W_d^{2,s}(\mathcal{X};R)) \leq \frac{c_2}{b}.
 	\end{equation}
@@
 
-\section{\textcolor{red}{In Progress Proofs}}
-
-\subsection{Proof of Theorem~\ref{thm:higher_order_sobolev_testing_rate}}
-
-Let $G = (V,E)$ be a graph over vertices $V = \set{v_1,\ldots,v_n}$, and let $\beta = (\beta_1,\ldots,\beta_n) \in \Reals^n$ be a signal over the vertices $V$.
-We observe responses $Y = (y_1,\ldots,y_n)$ according to the model
-\begin{equation*}
-y_i = \beta_i + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
-\end{equation*}
-Letting $L = VSV^T$ be the spectral decomposition of the Laplacian $L$ of $G$, our graph spectral test statistic is 
-\begin{equation*}
-T_{\spec} = \frac{1}{n}\sum_{k = 1}^{\kappa} \left(\sum_{i = 1}^{n} v_i y_i\right)^2
-\end{equation*}
-where $\kappa$ is a tuning parameter. The resulting test we will use is
-\begin{equation*}
-\phi_{\spec} = \1\{T_{\spec} \geq \frac{\kappa}{n} + t(b)\},~~\textrm{where}~ t(b) = b\sqrt{\frac{2\kappa}{n^2}}~~\textrm{for $b \geq 1$.}
-\end{equation*}
-
-Let $S_s(\beta;G)$ be a measure of smoothness the signal $\beta$ displays over the graph $G$, given by
-\begin{equation*}
-S_s(\beta;G) := \beta^T L^s \beta
-\end{equation*}
-In Lemma~\ref{lem:higher_order_fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}$.
-Our bound on the Type II error will be stated as a function of $S_2(\beta;G)$ as well as the $\kappa$th eigenvalue $s_{\kappa}$.
-
-\begin{lemma}
-	\label{lem:higher_order_fixed_graph_testing}
-	Let $1 \leq \kappa \leq n$ be an integer.
-	\begin{enumerate}
-		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}$ is upper bounded
-		\begin{equation}
-		\label{eqn:higher_order_graph_spectral_type_I_error}
-		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{1}{b^2}.
-		\end{equation}
-		\item \textbf{Type II error:} For any $b$ and $\beta$ such that
-		\begin{equation}
-		\label{eqn:higher_order_fixed_graph_testing_critical_radius}
-		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_s(\beta;G)}{ns_{\kappa}^s}
-		\end{equation}
-		the Type II error of $\phi_{\spec}$ is upper bounded,
-		\begin{equation}
-		\label{eqn:higher_order_graph_spectral_type_II_error}
-		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{3}{b}.
-		\end{equation}
-	\end{enumerate}
-\end{lemma}
-
-To prove Lemma~\ref{lem:higher_order_fixed_graph_testing} we will first compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}$, and then use Chebyshev's inequality to show \eqref{eqn:higher_order_graph_spectral_type_I_error} and \eqref{eqn:higher_order_graph_spectral_type_II_error}.
-
-\vspace{.2 in}
-
-\textit{Mean of $T_{\mathrm{spec}}$:} Using the notation $\dotp{v}{w} = \sum_{i = 1}^{n} v_iw_i$, we have
-\begin{align*}
-\Ebb(T_{\spec}) & = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{\beta}{v_k}^2 + \Ebb\bigl( \dotp{\varepsilon}{v_k}^2 + 2 \dotp{\varepsilon}{v_k} \dotp{\beta}{v_k}\bigr)\right) \\
-& = \frac{\kappa}{n} + \frac{1}{n}\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2.
-\end{align*}
-When $\beta = 0$, this equals $\kappa/n$. Otherwise, we have the following lower bound:
-\begin{align*}
-\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2 & = \norm{\beta}_2^2 - \sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \\
-& \geq \norm{\beta}_2^2 - \frac{1}{s_{\kappa}^s}\sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 s_k^s \\
-& \geq \norm{\beta}_2^2 - \frac{S_s(\beta;G)}{s_{\kappa}^s},
-\end{align*}
-and therefore $\Ebb(T_{\spec}) \geq \kappa/n + n^{-1}(\norm{\beta}_2^2 - S_s(\beta;G)/s_{\kappa}^s)$.
-
-\vspace{.2 in}
-
-\textit{Variance of $T_{\mathrm{spec}}$:}
-We write $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvectors $v_1,\ldots,v_{\kappa}$ as columns. Consequently,
-\begin{align}
-\Var(T_{\spec}) & = \frac{1}{n^2} \Var(y^T V_{\kappa} V_{\kappa}^T y) \\
-& = \frac{1}{n^2} \Var((\beta + \varepsilon)^T V_{\kappa} V_{\kappa}^T (\beta + \varepsilon)) \\
-& = \frac{1}{n^2} \Var(2 \beta^T V_{\kappa} V_{\kappa}^T \varepsilon + \varepsilon^T V_{\kappa} V_{\kappa}^T \varepsilon) \\
-& \leq \frac{1}{n^2}(4 \beta^T V_{\kappa} V_{\kappa}^T \beta + 2\kappa)
-\end{align}
-where the last inequality follows from standard properties of the Gaussian distribution.
-We now move on to showing the desired inequalities \eqref{eqn:higher_order_graph_spectral_type_I_error} and \eqref{eqn:higher_order_graph_spectral_type_II_error}.
-
-\vspace{.2 in}
-
-\textit{Proof of~\eqref{eqn:higher_order_graph_spectral_type_I_error}:} By Chebyshev's inequality,
-\begin{align*}
-\Pbb_{\beta = 0}\bigl(T_{\spec} \geq \frac{\kappa}{n} + t(b)\bigr)
-& \leq \Pbb_{\beta = 0}\bigl(\abs{T_{\spec} - \frac{\kappa}{n}} \geq t(b)\bigr) \\
-& \leq \frac{\Var_{\beta = 0}(T_{\spec})}{t(b)^2} = \frac{1}{b^2}.
-\end{align*}
-
-\vspace{.2 in}
-
-\textit{Proof of~\eqref{eqn:higher_order_graph_spectral_type_II_error}:} For simplicity, we introduce the notation
-\begin{equation*}
-\Delta = \frac{\norm{\beta}_2^2}{n} - \frac{S_s(\beta;G)}{ns_{\kappa}^s}.
-\end{equation*}
-Assumption~\eqref{eqn:higher_order_fixed_graph_testing_critical_radius} implies $\Delta \geq 2 t(b)$. Then another application of Chebyshev's inequality gives us
-\begin{align*}
-\Pbb_{\beta}\bigl(T_{\spec} \leq \frac{\kappa}{n} + t(b)\bigr) & = \Pbb_{\beta}\bigl(T_{\spec} - \Ebb_{\beta}(T_{\spec}) \leq t(b) - \Delta \bigr) \\
-& \leq \Pbb_{\beta}\bigl(\abs{T_{\spec} - \Ebb_{\beta}(T_{\spec})} \leq \Delta - t(b) \bigr) \tag{since $\Delta \geq t(b)$}	\\
-& \leq \frac{\Var_{\beta}(T_{\spec})}{(\Delta - t(b))^2} \\
-& \leq 4\frac{\Var_{\beta}(T_{\spec})}{\Delta^2} \tag{since $\Delta \geq 2t(b)$} \\
-& \leq 4\frac{2\kappa/n^2 + 4\beta^T V_{\kappa} V_{\kappa}^T \beta /n^2}{\Delta^2}.
-\end{align*}
-We handle each summand in the numerator separately. For the first term, since $\Delta \geq 2 t(b)$, we have
-\begin{equation}
-\label{eqn:higher_order_spectral_type_II_error_pf1}
-\frac{2\kappa}{n^2\Delta^2} \leq \frac{1}{2b^2}.
-\end{equation}
-
-For the second term, noting that $\Delta = \beta^T V_{\kappa} V_{\kappa}^T \beta/n$, we have
-\begin{align}
-\frac{\beta^T V_{\kappa} V_{\kappa}^T \beta/n^2}{\Delta^2} & = \frac{1}{n\Delta} \nonumber \\
-& \leq \frac{1}{2nt(b)} \nonumber \\
-& = \frac{1}{2b\sqrt{2\kappa}}, \label{eqn:higher_order_spectral_type_II_error_pf2}
-\end{align}
-and combining~\eqref{eqn:higher_order_spectral_type_II_error_pf1} and~\eqref{eqn:higher_order_spectral_type_II_error_pf2} yields~\eqref{eqn:higher_order_graph_spectral_type_II_error}.
-
-\subsubsection{Step 2: Bounding neighborhood graph functionals}
-
-To make use of Lemma~\ref{lem:fixed_graph_testing} we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$.
-In particular, we will show that for some constants $c_1,c_2,c_3$, the following statements:
-\begin{enumerate}[label=(E\arabic*)]
-	\item 
-	\label{event:higher_order_discrete_sobolev_norm}
-	\textbf{Graph Sobolev norm:} When $\beta = (f(x_1),\ldots,f(x_n))$ for $f \in W^{s,2}(\mathcal{X};R)$, and $r = n^{-1/(2 + d)}$,
-	\begin{equation}
-	\label{eqn:higher_order_continuous_to_discrete_sobolev_norm}
-	S_s(\beta;G_{n,r}) \leq c_1 \cdot n^{2 + s} r^{s(d + 2)} 
-	\end{equation}
-	\item 
-	\label{event:higher_order_eigenvalue_tail_decay}
-	\textbf{Eigenvalue tail bound:} For any $r \geq n^{-1/(2 + d)}$ and for all $\kappa \in [n]$:
-	\begin{equation}
-	\label{eqn:higher_ordereigenvalue_tail_bound}
-	s_{\kappa} \geq c_2 \cdot n r^{d + 2} \kappa^{2/d}
-	\end{equation}
-	\item 
-	\label{event:higher_order_l2_norm}
-	\textbf{Empirical norm of $f$:} When $f \in W^{s,2}(\mathcal{X};R)$ and \textcolor{red}{some other condition}
-	\begin{equation}
-	\label{eqn:higher_order_l2_to_empirical_norm}
-	\frac{1}{n}\sum_{i = 1}^{n}\beta_i^2 = \norm{f}_n^2 \geq c_3 \cdot \norm{f}_{\Leb^2}^2
-	\end{equation}
-\end{enumerate}
-
-each hold with probability $1 - o(1)$ as $n \to \infty$.
+Once proved, Theorem~\ref{thm:higher_order_sobolev_testing_rate} will give a unified treatment for all Sobolev spaces. To prove Theorem~\ref{thm:higher_order_sobolev_testing_rate}.
 
@@
 each hold with probability $1 - o(1)$ as $n \to \infty$.
+
+\textcolor{red}{TODO:} The inequality~\eqref{eqn:l2_to_empirical_norm} does not hold with probability $1 - o(1)$ but rather with probability $1 - c/b$ for an appropriate choice of constant $c$.
+Rewrite it to reflect this fact.
 
@@
 
+\subsection{Proof of Theorem~\ref{thm:higher_order_holder_testing_rate}}
+
+Let $G = (V,E)$ be a graph over vertices $V = \set{v_1,\ldots,v_n}$, and let $\beta = (\beta_1,\ldots,\beta_n) \in \Reals^n$ be a signal over the vertices $V$.
+We observe responses $Y = (y_1,\ldots,y_n)$ according to the model
+\begin{equation*}
+y_i = \beta_i + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
+\end{equation*}
+Letting $L = VSV^T$ be the spectral decomposition of the Laplacian $L$ of $G$, our graph spectral test statistic is 
+\begin{equation*}
+T_{\spec} = \frac{1}{n}\sum_{k = 1}^{\kappa} \left(\sum_{i = 1}^{n} v_i y_i\right)^2
+\end{equation*}
+where $\kappa$ is a tuning parameter. The resulting test we will use is
+\begin{equation*}
+\phi_{\spec} = \1\{T_{\spec} \geq \frac{\kappa}{n} + t(b)\},~~\textrm{where}~ t(b) = b\sqrt{\frac{2\kappa}{n^2}}~~\textrm{for $b \geq 1$.}
+\end{equation*}
+
+Let $S_s(\beta;G)$ be a measure of smoothness the signal $\beta$ displays over the graph $G$, given by
+\begin{equation*}
+S_s(\beta;G) := \beta^T L^s \beta
+\end{equation*}
+In Lemma~\ref{lem:higher_order_fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}$.
+Our bound on the Type II error will be stated as a function of $S_2(\beta;G)$ as well as the $\kappa$th eigenvalue $s_{\kappa}$.
+
+\begin{lemma}
+	\label{lem:higher_order_fixed_graph_testing}
+	Let $1 \leq \kappa \leq n$ be an integer.
+	\begin{enumerate}
+		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}$ is upper bounded
+		\begin{equation}
+		\label{eqn:higher_order_graph_spectral_type_I_error}
+		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{1}{b^2}.
+		\end{equation}
+		\item \textbf{Type II error:} For any $b$ and $\beta$ such that
+		\begin{equation}
+		\label{eqn:higher_order_fixed_graph_testing_critical_radius}
+		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_s(\beta;G)}{ns_{\kappa}^s}
+		\end{equation}
+		the Type II error of $\phi_{\spec}$ is upper bounded,
+		\begin{equation}
+		\label{eqn:higher_order_graph_spectral_type_II_error}
+		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{3}{b}.
+		\end{equation}
+	\end{enumerate}
+\end{lemma}
+
+To prove Lemma~\ref{lem:higher_order_fixed_graph_testing} we will first compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}$, and then use Chebyshev's inequality to show \eqref{eqn:higher_order_graph_spectral_type_I_error} and \eqref{eqn:higher_order_graph_spectral_type_II_error}.
+
+\vspace{.2 in}
+
+\textit{Mean of $T_{\mathrm{spec}}$:} Using the notation $\dotp{v}{w} = \sum_{i = 1}^{n} v_iw_i$, we have
+\begin{align*}
+\Ebb(T_{\spec}) & = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{\beta}{v_k}^2 + \Ebb\bigl( \dotp{\varepsilon}{v_k}^2 + 2 \dotp{\varepsilon}{v_k} \dotp{\beta}{v_k}\bigr)\right) \\
+& = \frac{\kappa}{n} + \frac{1}{n}\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2.
+\end{align*}
+When $\beta = 0$, this equals $\kappa/n$. Otherwise, we have the following lower bound:
+\begin{align*}
+\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2 & = \norm{\beta}_2^2 - \sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \\
+& \geq \norm{\beta}_2^2 - \frac{1}{s_{\kappa}^s}\sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 s_k^s \\
+& \geq \norm{\beta}_2^2 - \frac{S_s(\beta;G)}{s_{\kappa}^s},
+\end{align*}
+and therefore $\Ebb(T_{\spec}) \geq \kappa/n + n^{-1}(\norm{\beta}_2^2 - S_s(\beta;G)/s_{\kappa}^s)$.
+
+\vspace{.2 in}
+
+\textit{Variance of $T_{\mathrm{spec}}$:}
+We write $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvectors $v_1,\ldots,v_{\kappa}$ as columns. Consequently,
+\begin{align}
+\Var(T_{\spec}) & = \frac{1}{n^2} \Var(y^T V_{\kappa} V_{\kappa}^T y) \\
+& = \frac{1}{n^2} \Var((\beta + \varepsilon)^T V_{\kappa} V_{\kappa}^T (\beta + \varepsilon)) \\
+& = \frac{1}{n^2} \Var(2 \beta^T V_{\kappa} V_{\kappa}^T \varepsilon + \varepsilon^T V_{\kappa} V_{\kappa}^T \varepsilon) \\
+& \leq \frac{1}{n^2}(4 \beta^T V_{\kappa} V_{\kappa}^T \beta + 2\kappa)
+\end{align}
+where the last inequality follows from standard properties of the Gaussian distribution.
+We now move on to showing the desired inequalities \eqref{eqn:higher_order_graph_spectral_type_I_error} and \eqref{eqn:higher_order_graph_spectral_type_II_error}.
+
+\vspace{.2 in}
+
+\textit{Proof of~\eqref{eqn:higher_order_graph_spectral_type_I_error}:} By Chebyshev's inequality,
+\begin{align*}
+\Pbb_{\beta = 0}\bigl(T_{\spec} \geq \frac{\kappa}{n} + t(b)\bigr)
+& \leq \Pbb_{\beta = 0}\bigl(\abs{T_{\spec} - \frac{\kappa}{n}} \geq t(b)\bigr) \\
+& \leq \frac{\Var_{\beta = 0}(T_{\spec})}{t(b)^2} = \frac{1}{b^2}.
+\end{align*}
+
+\vspace{.2 in}
+
+\textit{Proof of~\eqref{eqn:higher_order_graph_spectral_type_II_error}:} For simplicity, we introduce the notation
+\begin{equation*}
+\Delta = \frac{\norm{\beta}_2^2}{n} - \frac{S_s(\beta;G)}{ns_{\kappa}^s}.
+\end{equation*}
+Assumption~\eqref{eqn:higher_order_fixed_graph_testing_critical_radius} implies $\Delta \geq 2 t(b)$. Then another application of Chebyshev's inequality gives us
+\begin{align*}
+\Pbb_{\beta}\bigl(T_{\spec} \leq \frac{\kappa}{n} + t(b)\bigr) & = \Pbb_{\beta}\bigl(T_{\spec} - \Ebb_{\beta}(T_{\spec}) \leq t(b) - \Delta \bigr) \\
+& \leq \Pbb_{\beta}\bigl(\abs{T_{\spec} - \Ebb_{\beta}(T_{\spec})} \leq \Delta - t(b) \bigr) \tag{since $\Delta \geq t(b)$}	\\
+& \leq \frac{\Var_{\beta}(T_{\spec})}{(\Delta - t(b))^2} \\
+& \leq 4\frac{\Var_{\beta}(T_{\spec})}{\Delta^2} \tag{since $\Delta \geq 2t(b)$} \\
+& \leq 4\frac{2\kappa/n^2 + 4\beta^T V_{\kappa} V_{\kappa}^T \beta /n^2}{\Delta^2}.
+\end{align*}
+We handle each summand in the numerator separately. For the first term, since $\Delta \geq 2 t(b)$, we have
+\begin{equation}
+\label{eqn:higher_order_spectral_type_II_error_pf1}
+\frac{2\kappa}{n^2\Delta^2} \leq \frac{1}{2b^2}.
+\end{equation}
+
+For the second term, noting that $\Delta = \beta^T V_{\kappa} V_{\kappa}^T \beta/n$, we have
+\begin{align}
+\frac{\beta^T V_{\kappa} V_{\kappa}^T \beta/n^2}{\Delta^2} & = \frac{1}{n\Delta} \nonumber \\
+& \leq \frac{1}{2nt(b)} \nonumber \\
+& = \frac{1}{2b\sqrt{2\kappa}}, \label{eqn:higher_order_spectral_type_II_error_pf2}
+\end{align}
+and combining~\eqref{eqn:higher_order_spectral_type_II_error_pf1} and~\eqref{eqn:higher_order_spectral_type_II_error_pf2} yields~\eqref{eqn:higher_order_graph_spectral_type_II_error}.
+
+\subsubsection{Step 2: Bounding neighborhood graph functionals}
+
+To make use of Lemma~\ref{lem:fixed_graph_testing} we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$.
+In particular, we will show that for some constants $c_1,c_2,c_3,c_4$ which may depend on $L$, $d$ and $s$ but do not depend on $n$, $f$ or $b$, the following statements:
+\begin{enumerate}[label=(E\arabic*)]
+	\item 
+	\label{event:higher_order_discrete_sobolev_norm}
+	\textbf{Graph Sobolev norm:} For $f \in C_d^{s}(\mathcal{X};R)$, $p \in C_d^{s - 1}(\Xset;R)$, and $1 \geq r(n) \geq n^{-1/(2(s - 1) + d)}$,
+	\begin{equation}
+	\label{eqn:higher_order_continuous_to_discrete_sobolev_norm}
+	S_s(f;G_{n,r}) \leq c_1 \cdot b \cdot n^{s + 1} r^{s(d + 2)} 
+	\end{equation}
+	\item 
+	\label{event:higher_order_eigenvalue_tail_decay}
+	\textbf{Eigenvalue tail bound:} For any $r \leq n^{-4/((2+d)(4s + d))}$ and for $\kappa = n^{2d/(4s + d)}$, 
+	\begin{equation}
+	\label{eqn:higher_order_eigenvalue_tail_bound}
+	s_{\kappa} \geq c_2 \cdot n r^{d + 2} \kappa^{2/d}
+	\end{equation}
+	\item 
+	\label{event:higher_order_l2_norm}
+	\textbf{Empirical norm of $f$:} When $f \in C^{s}(\mathcal{X};R)$ and $\norm{f}_{\Leb^2} \geq c_3 \cdot b \cdot n^{-4s/(4s + d)}$,
+	\begin{equation}
+	\label{eqn:higher_order_l2_to_empirical_norm}
+	\norm{f}_n^2 \geq \frac{1}{b} \cdot \norm{f}_{\Leb^2}^2
+	\end{equation}
+\end{enumerate}
+
+each hold with probability at least $1 - c_4/b$ for sufficiently large $n$.
+
+\paragraph{Proof of~\eqref{eqn:higher_order_continuous_to_discrete_sobolev_norm}:}
+
+We will take $s$ to be even, as the proof when $s$ is odd follows essentially the same steps. To simplify exposition, we introduce the iterated difference operator, defined recursively as
+\begin{equation*}
+D_{jk}f(x) = (D_{k}f(x_j) - D_{k}f(x))\frac{K_r(x_j,x)}{r^d}~~\textrm{for $j \in [n], k \in [n]^q$}, \quad D_jf(x) = (f(x_j) - f(x))\frac{K_r(x_j,x)}{r^d}
+\end{equation*}
+
+Now when $s$ is even, letting $q = s/2$ we have the decomposition
+\begin{equation}
+\label{eqn:higher_order_continuous_to_discrete_sobolev_norm_pf1}
+f^T L^s f =  \sum_{i = 1}^{n} \sum_{k \in [n]^q} \sum_{\ell \in [n]^q} r^{ds} D_kf(x_i) D_{\ell}f(x_i)
+\end{equation}
+For given index vectors $k,\ell$ and index $i$, let $I = \abs{k \cup \ell \cup i}$ be the total number of unique indices.
+We separate our analysis into cases based on the magnitude of $I$, specifically whether $I < s + 1$ or $I = s + 1$, and show that
+\begin{equation}
+\label{eqn:expected_difference_operators_1}
+\Ebb(D_kf(x_i) D_\ell f(x_i)) =
+\begin{cases*}
+O(r^{2s}), & ~~\textrm{if $I = s + 1$} \\
+O(r^{2} r^{d(I - (2q + 1))}), & ~~\textrm{otherwise}~ 
+\end{cases*}
+\end{equation}
+uniformly over $f \in C^{s}(L)$. 
+Before proving~\eqref{eqn:expected_difference_operators_1}, we verify that~\eqref{eqn:higher_order_continuous_to_discrete_sobolev_norm} is directly implied by~\eqref{eqn:expected_difference_operators_1}.
+In the sum on the right hand side of~\eqref{eqn:higher_order_continuous_to_discrete_sobolev_norm_pf1}, there are $O(n^{I})$ terms with exactly $I$ distinct indices.
+When $I < s + 1$, the total contribution of such terms to the sum is $O(n^{I}r^{d(I - 1) + 2})$.
+Since $r(n) \geq n^{-1/d}$, this increases with $I$.
+Taking $I = s$ to be the largest integer less than $s + 1$, the contribution of these terms to the sum is therefore $O(n^sr^{d(s - 1) + 2})$ which in light of the restriction $r \geq n^{-1/(2(s - 1) + d)}$ is $O(n^{s+1}r^{s(d +2)})$.
+On the other hand when $I = s + 1$, by~~\eqref{eqn:expected_difference_operators_1} we immediately have that the total contribution of these terms is $O(n^{s + 1}r^{2(s + d)})$.
+Therefore,
+\begin{equation*}
+\Ebb(f^T L^s f) = O(n^{s+1}r^{s(d+2)})
+\end{equation*}
+uniformly over $f \in C^s(L)$, and~\eqref{eqn:higher_order_continuous_to_discrete_sobolev_norm} by Markov's inequality.
+
+Now we prove~\eqref{eqn:expected_difference_operators_1}.
+Since $f \in C_d^s(R) \subseteq C_d^1(R)$, using a first-order Taylor expansion of $f(x)$ we can show that for all index vectors $k,\ell \in [n]^q$ and indices $i \in [n]$, the product of iterated difference operators $D_kf(x_i) D_{\ell}f(x_i)$ satisfies
+\begin{equation*}
+\abs{D_kf(x_i) D_{\ell}f(x_i)} \leq 4^{q} R^2 r^{2 - 2dq}
+\end{equation*}
+Moreover $D_kf(x_i) D_{\ell}f(x_i)$ will equal zero if there exists $x_j, j \in k \cup \ell \cup i$ such that
+\begin{equation*}
+\norm{x_j - x_h} > r,~~\textrm{for all $h \neq j \in k \cup \ell \cup i$}
+\end{equation*}
+Therefore $D_kf(x_i) D_{\ell}f(x_i)$ is nonzero with probability $O(r^{d(\abs{k \cup \ell \cup i} - 1)})$, which along with the boundedness of $D_kf(x_i) D_{\ell}f(x_i)$ implies the second upper bound in~\eqref{eqn:expected_difference_operators_1}.
+
+To show the first upper bound in~\eqref{eqn:expected_difference_operators_1}, we apply the law of iterated expectation, 
+\begin{align}
+\Ebb\left[D_kf(x_i)D_{\ell}f(x_i)\right] & = \Ebb\left[\Ebb\left(D_kf(x)|x_i = x\right) \Ebb\left(D_{\ell}f(x)|x_i = x\right)\right] \nonumber \\
+& = \Ebb\left[\Ebb\left(D_kf(x)|x_i = x\right)^2\right]. \label{eqn:expected_difference_operators_pf1}
+\end{align}
+By Lemma~\ref{lem:leading_term}, we have that $\Ebb(D_kf(x)) = O(r^s)$ for all values of $x$, implying the desired result.
+
+\paragraph{Proof of~\eqref{eqn:higher_order_eigenvalue_tail_bound}:}
+
+We prove~\eqref{eqn:higher_order_eigenvalue_tail_bound} by comparing $G_{n,r}$ to the tensor product of a $d$-dimensional lattice and a complete graph.
+The latter is a highly structured graph with known eigenvalues, which as we will see are sufficiently lower bounded for our purposes.
+
+Let $\wt{r} = r/(3(\sqrt{d} + 1)), M = (1/\wt{r})^d, N = n\wt{r}^d$. Assume without loss of generality that $M$ and $N$ are integers. Additionally, for $t = n^{1/d}$ and $m = M^{1/d}$ let 
+\begin{equation*}
+\overline{X} = \set{\frac{1}{t}(k_1,\ldots,k_d): k \in [t]^d},~~ \overline{Z} = \set{\frac{1}{m}(j_1,\ldots,j_d): j \in [m]^d}.
+\end{equation*}
+For a given $\overline{z}_j \in \overline{Z}$, we write $Q(z_j) = m^{-1}[j_1 - 1,j_1] \times \cdots \times m^{-1}[j_d - 1,j_d]$ for the cube of side length $1/m$ with $z_j$ at one corner.
+
+Consider the graph $H = (\overline{X}, E_H)$, where $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$ if
+\begin{equation*}
+\textrm{there exists}~\ol{z}_i, \ol{z}_j \in \ol{Z}~\textrm{such that}~\ol{x}_k \in Q(\ol{z}_i),~ \ol{x}_\ell \in Q(\ol{z}_j),~\textrm{and}~\norm{i - j}_1 \leq 1.
+\end{equation*}
+On the one hand $H \cong \ol{G}^M_d \otimes K_N$ where $\ol{G}^M_d$ is the $d$-dimensional lattice on $M$ nodes, and $K_N$ is the complete graph on $N$ nodes.
+On the other hand, we now show that with high probability $G_{n,r} \succeq H$.
+If $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$, then there exist $\ol{z}_i, \ol{z}_j$ such that
+\begin{equation*}
+\norm{\ol{x}_k - \ol{x}_{\ell}}_2 \leq m^{-1} + \norm{\ol{x}_k - \ol{z}_{i}}_2 + \norm{\ol{x}_{\ell} - \ol{z}_{j}}_2 \leq \wt{r}(1 + \sqrt{d}) = r/3.
+\end{equation*}
+
+Assuming~\eqref{eqn:transport_distance} holds, if $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$, then for sufficiently large $n$
+\begin{equation*}
+\norm{\pi(\ol{x}_k) - \pi(\ol{x}_\ell)}_2 \leq 2 c \left(\frac{\log n}{n}\right)^{1/d} + \frac{r}{3} \leq r,
+\end{equation*}
+implying that $(\pi(\ol{x}_k), \pi(\ol{x}_{\ell})) \in E$. Therefore, $G_{n,r} \succeq \ol{G}^M_d \otimes K_N$ whenever~\eqref{eqn:transport_distance} holds.
+
+The eigenvalues of lattices and complete graphs are known to satisfy, respectively
+\begin{equation*}
+\lambda_k(\ol{G}^{M}_d) \geq \frac{k^{2/d}}{M^{2/d}}~\textrm{for $k = 0,\ldots,M - 1$},~~ \textrm{and}~\lambda_{j}(K_N) \geq N\1\{j > 0\}~\textrm{for $j = 0,\ldots,N-1$.}
+\end{equation*}
+and by standard facts regarding the eigenvalues of tensor product graphs, we have that the spectrum $\Lambda(H)$ satisfies
+\begin{equation*}
+\Lambda(H) = \set{N\lambda_k(\ol{G}^{M}_d) + M\lambda_j(K_N): \textrm{for $k = 0,\ldots,M - 1$ and $j = 0,\ldots,N-1$}}
+\end{equation*}
+For all $j = 1,\ldots,N-1$, we have that $M\lambda_j(K_N) = MN = n$. Therefore,
+\begin{align*}
+\lambda_{\kappa}(H) & \geq \{n \wedge N\lambda_{\kappa}(\ol{G}^{M}_d)\} \\
+& \geq \{n \wedge n\wt{r}^d\frac{\kappa^{2/d}}{M^{2/d}}\} \\
+& \geq \{n \wedge (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d}\} \\
+& \geq (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d},
+\end{align*}
+where the last inequality can be verified by a quick calculation in light of $\kappa = n^{2d/(4s + d)}$ and $r \leq n^{-4/((2+d)(4s + d))}$.
+Since we've already shown that $\lambda_{\kappa}(G_{n,r}) \geq \lambda_{\kappa}(H)$ when~\eqref{eqn:transport_distance} is satisfied, which happens probability $1 - o(n^{-1})$, this completes the proof of~\eqref{eqn:higher_order_eigenvalue_tail_bound}.
+
+\paragraph{Proof of~\eqref{eqn:higher_order_l2_to_empirical_norm}:}
+
+Let $Z = \frac{1}{n}\sum_{i = 1}^{n} f^2(x_i)$. We upper bound $\Ebb[Z^2]$,
+\begin{align*}
+\Ebb[Z^2] & \leq \Ebb(f^2(x_1))^2 + \frac{1}{n}\Ebb(f^4(x_1)) \\
+& \leq \Ebb(f^2(x_1))^2 + \frac{R^2}{n}\Ebb(f^2(x_1)) \\
+& \overset{(i)}{\leq} \Ebb(f^2(x_1))^2\left(1 + \frac{R^2}{n/\Ebb(f^2(x_1))}\right) \\
+& \overset{(ii)}{\leq} \Ebb(f^2(x_1))^2\left(1 + \frac{R^2}{c_3^2 b^2 n^{d/(4s + d)}}\right)
+\end{align*}
+where $(i)$ follows since $f \in C_d^s(\Xset;R)$ implies $\abs{f(x)} \leq R$, and $(ii)$ follows by assumption. The statement then follows by Lemma~\ref{lem:paley_zygmund}.
+
+\subsubsection{Step 3: Conclusion}
+
+We note that for all possible values of $X \in \Xset^n$, under the null hypothesis $f = f_0 = 0$ and therefore $\beta = (f(x_1),\ldots,f(x_n)) = 0$ as well.
+Therefore by~\eqref{eqn:higher_order_graph_spectral_type_I_error}, we have the following bound on Type I error:
+\begin{equation}
+\Ebb_{f_0}(\phi_{\mathrm{spec}}) = \mathbb{E}(\mathbb{E}_{\beta = 0}(\phi_{\spec}) | X) \leq \frac{1}{b^2}.
+\end{equation}
+
+Now, we bound Type II error under the assumption $f \in C_d^s(\mathcal{X};R), p \in C_d^{s - 1}(\Xset;R)$ uniformly bounded away from $0$ and $\infty$ over $\Xset$, and 
+\begin{equation}
+\label{eqn:higher_order_critical_radius_1}
+\norm{f}_{\Leb^2}^2 \geq \epsilon^2 = c_3^2 \cdot b^2 \cdot n^{-4s/(4s + d)}.
+\end{equation}
+Choosing $n^{-1/(2(s - 1) + d)}\leq r(n) \leq n^{-4/((2+d)(4s + d))}$, we may therefore apply our conclusions in Step 2; namely, that for every possible choice of $f$ there exists a good set $\mathcal{E}_f \subseteq \Xset^n$ with $\Pbb(\mathcal{E}_f) \geq 1 - c_4/b$ such that each of \eqref{eqn:continuous_to_discrete_sobolev_norm}, \eqref{eqn:eigenvalue_tail_bound}, and \eqref{eqn:l2_to_empirical_norm} hold for all $X \subseteq \mathcal{E}_f$.
+Choosing $\kappa = n^{2d/(4s + d)}$ to balance the squared bias and variance terms on the right hand side of~\eqref{eqn:higher_order_fixed_graph_testing_critical_radius}, we have that for all $X \subseteq \mathcal{E}_f$
+\begin{align*}
+2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_s(\beta;G_{n,r})}{ns_{\kappa}} & \leq 2bn^{-4s/(4s+d)} + c\cdot b \cdot \frac{n^{s} r^{s(d + 2)}}{s_{\kappa}^s} \tag{by \eqref{eqn:continuous_to_discrete_sobolev_norm}} \\
+& \leq 2bn^{-4s/(4s+d)} + c\cdot b \cdot\frac{1}{n^{4s/(4s+d)}} \tag{by \eqref{eqn:eigenvalue_tail_bound}} \\
+& \leq \frac{1}{b}\norm{f}_{\Leb^2} \\
+& \leq \frac{1}{n}\sum_{i = 1} \beta_i^2. \tag{by \eqref{eqn:l2_to_empirical_norm}}
+\end{align*}
+where the last two inequalities follow for a suitably large choice of $c_3$ in~\eqref{eqn:higher_order_critical_radius_1}.
+We conclude that for all $X \subseteq \mathcal{E}_f$, the inequality \eqref{eqn:higher_order_fixed_graph_testing_critical_radius} is satisfied with respect to $\beta = (f(x_1),\ldots,f(x_n))$ and $G = G_{n,r}$.
+As a result the worst-case Type II error is bounded
+\begin{equation*}
+\sup_{\substack{f \in C_d^{s}(\mathcal{X;R}), \\ \norm{f}_{\Leb^2} \geq \epsilon}}\mathbb{E}_{f}(1 - \phi_{\spec}) \leq \sup_{\substack{f \in C_d^{s}(\mathcal{X;R}), \\ \norm{f}_{\Leb^2} \geq \epsilon}} \mathbb{E}\bigl[\mathbb{E}_{\beta}(1 - \phi_{\spec}|X \in \mathcal{E}_f)\bigr] + \frac{c_4}{b} \leq \frac{3 + c_4}{b},
+\end{equation*}
+completing the proof of Theorem~\ref{thm:higher_order_holder_testing_rate}.
+
 \subsection{Proof of Theorem~\ref{thm:twosample_sobolev_testing_rate}}
@@
 \begin{lemma}
+	\label{lem:leading_term}
+	Let $s$ be a positive integer, and $R \geq 0$. Suppose $f \in C^s(R)$, $p \in C^{s-1}(R)$, $k \in [n]^q$ for some $q \geq 1$, and that $K_r$ is a $2$nd-order kernel. Then if $2q \leq s - 1$,
+	\begin{equation}
+	\label{eqn:leading_term}
+	\Ebb(D_kf(x)) = \sum_{\ell = 2q}^{s - 1} O(r^{\ell}) \cdot f_{\ell}(x) + O(r^s),
+	\end{equation}
+	for some $f_{\ell} \in C^{s - \ell}(L)$. If $2q \geq s$, $\Ebb(D_kf(x)) = O(r^s)$.  All $O(\cdot)$ terms may depend $L$ and $s$, but do not depend on $f$ or $x$.
+\end{lemma}
+\begin{proof}
+	We will prove Lemma~\ref{lem:leading_term} in the case where $d = 1$. When $d \geq 2$, using multivariate Taylor expansions we find the same result, but with more notational overhead.
+
+	We will prove by induction on $q$ in the case where $d = 1$. The When $q = 1$ and $k \in [n]$, by taking Taylor expansions of $f$ and $p$, we obtain
+	\begin{align}
+	\nonumber 
+	\Ebb(D_kf(x)) & = \sum_{\ell = 1}^{s - 1} f^{(\ell)}(x) \int (z - x)^{\ell} K_r(z,x) p(z) dz + O(r^s) \\ 
+	& = \sum_{\ell = 1}^{s - 1} \sum_{a = 0}^{s - 1} f^{(\ell)}(x) p^{(a)}(x) \underbrace{\int(z - x)^{\ell + a} K_r(z,x) dz}_{:=I_{t + a}} + O(r^s) \label{eqn:leading_term_pf1}
+	\end{align}
+	Since $K$ is a $2$nd-order kernel, $I_{1} = 0$ and $I_t = O(r^{t})$ for $t \geq 2$.
+Additionally, when $\ell + a \leq s - 1$, we have that $f^{(\ell)}p^{(a)} \in C^{\min\{s - \ell, s - 1 - a\}} \subseteq C^{s - (\ell + a)}$, and $\abs{f^{(\ell)}p^{(a)}} \leq L^2$ for any $\ell$ and $a$.
+We can therefore simplify~\eqref{eqn:leading_term_pf1} by combining all terms where $\ell + a = t$, obtaining
+	\begin{equation}
+	\label{eqn:leading_term_pf2}
+	\Ebb(D_kf(x)) = \sum_{t = 1}^{s - 1} f_t(x) I_t + O(r^s) = \sum_{t = 2}^{s - 1} f_t(x) I_t + O(r^s),
+	\end{equation}
+	which establishes~\eqref{eqn:leading_term} in the base case.
+
+	Now, we assume \eqref{eqn:leading_term} holds for all $k \in [n]^q$, and prove the desired estimate on $\Ebb(D_{kj}f(x))$ for each $j \in [n]$.
+By the law of iterated expectation and~\eqref{eqn:leading_term_pf2},
+	\begin{align*}
+	\Ebb(D_{kj}f(x)) & = \Ebb(D_k(\Ebb(D_jf))(x)) \\
+	& = \Ebb\left(D_k\left(\sum_{t = 2}^{s - 1} I_t f_t + O(r^s)\right)(x)\right) \\
+	& = \sum_{t = 2}^{s - 1} I_t \cdot \Ebb(D_kf_t(x)) + O(r^s)
+	\end{align*}
+	where the second equality follows from the linearity and boundedness of $f \mapsto \Ebb(D_kf)$.
+We now apply the inductive hypothesis to $\Ebb(D_kf_t(x))$.
+If $2(q + 1) \geq s$, note that since $f_t \in C^{s - t}(L)$ for $t \geq 2$, we have by hypothesis $\Ebb(D_kf_t(x)) = O(r^{s - t})$.
+As a result
+	\begin{equation*}
+	\sum_{t = 2}^{s - 1} I_t \cdot \Ebb(D_kf_t(x)) = \sum_{t = 2}^{s - 1} I_t \cdot O(r^{s - t}) = O(r^s)
+	\end{equation*} 
+	Otherwise $2(q + 1) \leq s - 1$.
+For each $t = 2,\ldots, s - 1$, if additionally  $2q \leq s - t - 1$, then by hypothesis $\Ebb(D_kf_t(x)) = \sum_{\ell = 2q}^{s - t - 1} O(r^{\ell}) \cdot g_{\ell}(x) + O(r^{s - t})$ for some $g_{\ell} \in C^{s - t - \ell}(L)$, and otherwise $\Ebb(D_kf_t(x)) = O(r^{s - t})$.
+Therefore,
+	\begin{align*}
+	\sum_{t = 2}^{s - 1} I_t \cdot \Ebb(D_kf_t(x)) & = \sum_{t = 2}^{s - 1 - 2q} I_t \cdot \left\{\sum_{\ell = 2q}^{s - t - 1} O(r^{\ell}) \cdot g_{\ell}(x) + O(r^{s - t}) \right\} + \sum_{t = s - 1 - 2q}^{s - 1}I_{t} \cdot O(r^{s - t}) \\
+	& = \sum_{t = 2}^{s - 1 - 2q} I_t \cdot \left\{\sum_{\ell = 2q}^{s - t - 1} O(r^{\ell}) \cdot g_{\ell}(x)\right\}  + O(r^s) \\
+	& = \sum_{\ell = 2q}^{s - 3} \sum_{t = 2}^{s - \ell - 1} I_t \cdot O(r^{\ell}) \cdot g_{\ell}(x) + O(r^s).
+	\end{align*}
+	Noting that $g_{\ell} \in C^{s - (t + \ell)}(L)$ for some $\ell + t = 2(q + 1),\ldots,s - 1$, and $I_t \cdot O(r^{\ell}) = O(r^{t + \ell})$, we can rewrite the final equation as a sum over $\ell + t = 2(q + 1),\ldots,s - 1$, which proves~\eqref{eqn:leading_term}.
+\end{proof}
+
+\begin{lemma}
 	\label{lem:paley_zygmund}

### 2020-01-13 notes/graph_testing_summary.tex
added minimax regression testing rate for higher order Holder classes

@@
 \end{equation*}
-The worst-case risk may be quite close to $1$ unless we require some separation between null and alternative spaces. A more realistic measure of performance is therefore
+The worst-case risk may be quite close to $1$ unless we enforce some separation between null and alternative spaces. A more realistic measure of performance is therefore
 \begin{equation*}
-\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, \norm{f - f_0}_2 \geq \epsilon} \Ebb_f(1 - \phi).
+\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, \norm{f - f_0}_{\Leb^2} \geq \epsilon} \Ebb_f(1 - \phi).
 \end{equation*}
@@
 	\label{eqn:sobolev_testing_rate_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; \mathcal{W}^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}.
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}.
 	\end{equation}
 \end{theorem}
-The testing rate~\eqref{eqn:sobolev_testing_rate} matches the minimax critical radius up to a factor of $\log^{h(a,d)}n$.
 
@@
 	Let $b \geq 1$ and $a > 0$ be fixed constants, and let $d = 1$.
-Suppose that $P$ and $Q$ are absolutely continuous probability measures over $\mathcal{X} = [0,1]$ with density functions $p$ and $q$ bounded above and below by constants, i.e
+Suppose that $\mu = (P + Q)/2$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]$ with density functions $\rho$ bounded above and below by constants, i.e
 	\begin{equation*}
-	0 < p_{\min},q_{\min} < p(x),q(x) < p_{\max},q_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
+	0 < \rho_{\min} < \rho(x) < \rho_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
 	\end{equation*}
@@
 
-For a given $s > 0$, the $s$th Holder norm is given by
+For a given $s > 0$, let $\ell = \floor{s}$ be the largest integer strictly less than $s$. Then the $s$th Holder norm is given by
 \begin{equation*}
-\norm{f}_{C_d^{s}(\mathcal{X})} := \sum_{\abs{\alpha} \leq s} \norm{D^{\alpha}f}_{\infty} + \sum_{\abs{\alpha} = s} \sup_{x,y \in \mathcal{X}} \frac{\abs{D^{\alpha}f(y) - D^{\alpha}f(x)}}{\norm{x - y}_2}
+\norm{f}_{C_d^{s}(\mathcal{X})} := \sum_{\abs{\alpha} < s} \norm{D^{\alpha}f}_{\infty} + \sum_{\abs{\alpha} = \ell} \sup_{x,y \in \mathcal{X}} \frac{\abs{D^{\alpha}f(y) - D^{\alpha}f(x)}}{\norm{x - y}_2^{s - \ell}}
 \end{equation*}
-and the $s$th Holder space $C_d^{s}(\mathcal{X})$ consists of all functions which are $s$ times continuously differentiable with finite $s$ Holder norm.
+and the $s$th Holder space $C_d^{s}(\mathcal{X})$ consists of all functions which are $\ell$ times continuously differentiable with finite Holder norm $\norm{\cdot}_{C_d^{s}(\mathcal{X})}$.
 Denote the Holder ball by $C_d^{s}(\mathcal{X},R) = \set{f \in C_d^{s}(\mathcal{X}): \norm{f}_{C_d^{s}(\mathcal{X})} \leq R}$.

### 2020-01-13 notes/weekly_notes/12.20.19.NotesForWeek.tex
added minimax regression testing rate for higher order Holder classes

@@
 \label{eqn:roughness_functional_bound}
-\mathbb{E}\left[R_{s,n}(f)\right] = O\left(L^2\right)
+\mathbb{E}\left[R_{s,n}(f)\right] = O\left(1\right)
 \end{equation}
-
+uniformly over all $f \in C^{s}(1)$. 
 These notes detail our current progress in showing such a result for different values of $s$.
@@
 \end{equation}
-We have already obtained the desired bound on the 1st-order roughness functional $R_{1,n}(f)$, for any $f \in W_d^{1}(L)$, all scalings $r = r(n)$, and the uniform kernel.
-We therefore begin these notes with the case $s = 2$.
 
-\section{Bounding the 2nd-Order Roughness Functional}
-The (first and) second order roughness functionals are somewhat special.
-These roughness functionals do not involve taking compositions of the difference operator $D_k$, and as a result, we get the desired bounds in expectation even when $r(n)$ is small.
-\begin{lemma}
-	\label{lem:2nd_order_roughness_functional}
-	Fix $L > 0$, and form the neighborhood graph $G_{n,r}$ using kernel $K(z) = \1\{z^2 \leq r^2\}$. Suppose $f \in C^2(L)$ and $p \in C^1(L)$. Then for any $r(n) \geq n^{-1/(2+d)}$ and , we have
+\section{Bounding the Roughness Functional in Expectation}
+
+For an integer $\ell > 0$, we will say $K$ is a $2$nd-order kernel if it is bounded and compactly supported on $B(0,1)$ and additionally
 \begin{equation*}
-	\Ebb(R_{2,n}(f)) = O(L^2)
+\int K(z) = 1 ~~\textrm{and}~~\int z K(z) = 0
 \end{equation*}
-\end{lemma}
-
-\section{Bounding the 3rd-Order Roughness Functional}
-
-For the order-$s$ roughness functionals when $s \geq 3$, the graph Laplacian is forced to approximate higher order derivatives by composing difference operators-- that is, taking differences of differences, etc.
-In for the roughness functional to continue scaling at the desired rate, we will need $r(n)$ to be larger, and we cannot use the uniform kernel $K$ any longer.
-In the particular case $s = 3$, we will require a kernel $K$ which is bounded and compactly supported on $B(0,1)$, and additionally satisfies
-\begin{equation*}
-\int z^j K(z) = 0~~\textrm{for $j = 1,2$.}
-\end{equation*}
-We then have the following result.
-\begin{lemma}
-	\label{lem:3rd_order_roughness_functional}
-	Suppose $f \in C^{3}(L)$ for some $L > 0$, and further suppose $p \in C^{2}(L)$. For any kernel $K$ which satisfies the previous conditions, and any $r(n) \geq n^{-1/(4+d)}$, we have that
-	\begin{equation*}
-	\Ebb(R_{s,n}(f)) = O(1)
-	\end{equation*} 
-\end{lemma}
-
-\section{\textcolor{red}{Bounding the Roughness Functional in Expectation}}
-
-\textcolor{red}{WARNING:} The following Lemma is not proved, but it is the result I am trying to prove. After the Lemma is stated, I go into some detail about how I intend to prove it.
-
-For an integer $\ell > 0$, we will say $K$ is an order-$\ell$ kernel
-\begin{equation*}
-\int K(z) = 1 ~~\textrm{and}~~\int z^j K(z) = 0,~~\textrm{for $j = 1,\ldots,\ell$}
-\end{equation*}
-and further $K$ is $\ell$ times continuously differentiable with $K^{j}(z) = O(z^j K(z))$ for $j = 1,\ldots,s$.
+That $K$ be a $2$nd-order kernel is crucial for $L_n$ to act as a $2$nd-order differential operator, and for $R_{s,n}$ to scale at the proper rate.
 
@@
 	\label{lem:roughness_functional_expectation}
-	Suppose $f \in C^{s}(L)$ for some $s,L > 0$, and further suppose $p \in C^{\ell}(L)$ for $\ell = \floor{s}$ the largest integer strictly less than $s$.
-For any order-$\ell$ kernel $K$ and any $1 \geq r(n) \geq n^{-1/(2(s - 1)+d)}$, we have that
+	Suppose $f \in C^{s}(L)$ for some positive integer $s$ and $L > 0$, and further suppose $p \in C^{\ell}(L)$ for $\ell = \floor{s}$ the largest integer strictly less than $s$.
+For any $2$nd-order kernel $K$ and any $1 \geq r(n) \geq n^{-1/(2(s - 1)+d)}$, we have that
 	\begin{equation*}
-	\Ebb(R_{s,n}(f)) = O(1)
+	\Ebb(R_{s,n}(f)) = O(1),
 	\end{equation*} 
+	uniformly over all $f \in C^{s}(L)$.
 \end{lemma}
-When $s \geq 3$ the graph Laplacian approximates higher-order derivatives by composing difference operators.
-The crudeness of this approximation means that in Lemma~\ref{lem:roughness_functional_expectation} we allow for a slower scaling of $r(n)$ than we did in Lemmas~\ref{lem:2nd_order_roughness_functional} and~\ref{lem:3rd_order_roughness_functional}.
-At a high level, however, we prove Lemma~\ref{lem:roughness_functional_expectation} in a similar fashion to Lemma~\ref{lem:2nd_order_roughness_functional}.
-We first break the sum into cases according to the number of unique indices, then bound the expectation case by case.
+
+Note that Lemma~\ref{lem:roughness_functional_expectation} implies a high-probability bound on $R_{s,n}(f)$ by Markov's inequality.
+To prove Lemma~\ref{lem:roughness_functional_expectation}, we first break the sum~\eqref{eqn:roughness_functional_representation_even} (or ~\eqref{eqn:roughness_functional_representation_odd}) into cases according to the number of unique indices, then bound the expectation case by case.
 The following Lemma will be the workhorse which supplies a sufficient bound in each case.
@@
 
+The minimum scaling condition $r(n) \geq n^{-1/(2(s - 1) + d)}$ is needed in order to ensure the diagonal terms of the sums in~\eqref{eqn:roughness_functional_representation_even} and~\eqref{eqn:roughness_functional_representation_odd} (where not all indices are distinct) do not dominate the off-diagonal terms (where all indices are distinct).
+
 \section{Proofs}
@@
 Since there are $O(n^{s+1})$ such terms, we have that the total contribution of these terms is $O(n^{s + 1}r^{2s}n^{-(s + 1)}r^{-2s}) = O(1)$.
+
+\section{Proof of Lemma~\ref{lem:expected_difference_operators}}
+
+\subsection{Proof of~\eqref{eqn:expected_difference_operators_1}}
+
+Let $k,\ell \in [n]^q$ be index vectors. We first prove the desired bound in the case when some indices are repeated, and then the desired bound in the case when all indices are distinct.
+
+\paragraph{Repeated indices.}
+
+Since $f \in C^1(L)$, for all index vectors $k,\ell \in [n]^q$ and indices $i \in [n]$, the product of iterated difference operators $D_kf(x_i) D_{\ell}f(x_i)$ satisfies
+\begin{equation*}
+\abs{D_kf(x_i) D_{\ell}f(x_i)} \leq 4^{q} L^2 r^{2 - ds}
+\end{equation*}
+Moreover $D_kf(x_i) D_{\ell}f(x_i)$ will equal zero if there exists $x_j, j \in k \cup \ell \cup i$ such that
+\begin{equation*}
+\norm{x_j - x_h} > r,~~\textrm{for all $h \in k \cup \ell \cup i$}
+\end{equation*}
+Therefore $D_kf(x_i) D_{\ell}f(x_i)$ is nonzero with probability $O(r^{d(\abs{k \cup \ell \cup i} - 1)})$, which along with the boundedness of $D_kf(x_i) D_{\ell}f(x_i)$ implies the claimed result.
+
+\paragraph{All indices distinct.}
+
+By the law of iterated expectation, we have
+\begin{align}
+\Ebb\left[D_kf(x_i)D_{\ell}f(x_i)\right] & = \Ebb\left[\Ebb\left(D_kf(x)|x_i = x\right) \Ebb\left(D_{\ell}f(x)|x_i = x\right)\right] \nonumber \\
+& = \Ebb\left[\Ebb\left(D_kf(x)|x_i = x\right)^2\right] \label{eqn:expected_difference_operators_pf1}
+\end{align}
+The result then follows from Lemma~\ref{lem:leading_term}.
+
+\subsection{Proof of~\eqref{eqn:expected_difference_operators_2}}
+
+The proof of~\eqref{eqn:expected_difference_operators_2} when some index is repeated is essentially the same, and we do not reproduce it.
+
+When all indices $\ell,k,i$ and $j$ are distinct, by Lemma~\ref{lem:leading_term} there exists some $f_{s-1} \in C^1(L)$ such that
+\begin{align*}
+\Ebb(d_iD_kf(x_j)d_iD_{\ell}f(x_j)K_r(x_i,x_j)) & = \Ebb\biggl(\Bigl(d_i\bigl(O(r^{s-1})\cdot f_{s - 1} + O(r^s)\bigr)\Bigr)^2K_r(x_i,x_j)\biggr) \\
+& = O(r^{2s - 2})\Ebb\Bigl(\bigl(d_if_{s-1}(x_j))\bigr)^2K_{r}(x_i,x_j)\Bigr) + O(r^{2s - 1})\Ebb\bigl(D_if_{s - 1}(x_j)\bigr)+ O(r^{2s})
+\end{align*}
+where the second equality follows from the linearity of the difference operator. Then since $f_{s-1} \in C^1(L)$, we have $\Ebb(D_if_{s-1}(x_j)) = O(r)$, and
+\begin{equation*}
+\Ebb\Bigl(\bigl(d_if_{s-1}(x_j))\bigr)^2K_{r}(x_i,x_j)\Bigr) = \Ebb\Bigl(\bigl(O(r)\bigr)^2K_{r}(x_i,x_j)\Bigr) = O(r^2),
+\end{equation*}
+establishing $\Ebb(d_iD_kf(x_j)d_iD_{\ell}f(x_j)K_r(x_i,x_j)) = O(r^{2s})$ as claimed.
+\section{Auxiliary Results.}
+\begin{lemma}
+	\label{lem:difference_operator_taylor}
+	When $f \in C^s(L)$, for any $k \in [n]$ 
+	\begin{equation*}
+	D_kf(x) = \left(\sum_{r = 1}^{s - 1} (x_k - x)^r f^{(r)}(x) + O((x_k - x)^{s})\right) \cdot K_r(x_k,x)
+	\end{equation*}
+\end{lemma}
+
+\begin{lemma}
+	\label{lem:expected_kernel}
+	When $K$ is a bounded and compactly supported kernel, then for any $s$ and for all $x \in \Rd$
+	\begin{equation*}
+	\Ebb(\abs{K(x_i,x)}^s) \leq CLr^{(1 - s)d}
+	\end{equation*}
+\end{lemma}
+
+\begin{lemma}
+	\label{lem:leading_term}
+	Suppose $f \in C^s(L)$, $p \in C^{s-1}(L)$, $k \in [n]^q$ for some $q \geq 1$, and that $K_r$ is a $2$nd-order kernel. Then if $2q \leq s - 1$,
+	\begin{equation}
+	\label{eqn:leading_term}
+	\Ebb(D_kf(x)) = \sum_{\ell = 2q}^{s - 1} O(r^{\ell}) \cdot f_{\ell}(x) + O(r^s),
+	\end{equation}
+	for some $f_{\ell} \in C^{s - \ell}(L)$. If $2q \geq s$, $\Ebb(D_kf(x)) = O(r^s)$.  All $O(\cdot)$ terms may depend $L$ and $s$, but do not depend on $f$ or $x$.
+\end{lemma}
+\begin{proof}
+	We will prove Lemma~\ref{lem:leading_term} in the case where $d = 1$. When $d \geq 2$, using multivariate Taylor expansions we find the same result, but with more notational overhead.
+
+	We will prove by induction on $q$ in the case where $d = 1$. The When $q = 1$ and $k \in [n]$, by taking Taylor expansions of $f$ and $p$, we obtain
+	\begin{align}
+	\nonumber 
+	\Ebb(D_kf(x)) & = \sum_{\ell = 1}^{s - 1} f^{(\ell)}(x) \int (z - x)^{\ell} K_r(z,x) p(z) dz + O(r^s) \\ 
+	& = \sum_{\ell = 1}^{s - 1} \sum_{a = 0}^{s - 1} f^{(\ell)}(x) p^{(a)}(x) \underbrace{\int(z - x)^{\ell + a} K_r(z,x) dz}_{:=I_{t + a}} + O(r^s) \label{eqn:leading_term_pf1}
+	\end{align}
+	Since $K$ is a $2$nd-order kernel, $I_{1} = 0$ and $I_t = O(r^{t})$ for $t \geq 2$.
+Additionally, when $\ell + a \leq s - 1$, we have that $f^{(\ell)}p^{(a)} \in C^{\min\{s - \ell, s - 1 - a\}} \subseteq C^{s - (\ell + a)}$, and $\abs{f^{(\ell)}p^{(a)}} \leq L^2$ for any $\ell$ and $a$.
+We can therefore simplify~\eqref{eqn:leading_term_pf1} by combining all terms where $\ell + a = t$, obtaining
+	\begin{equation}
+	\label{eqn:leading_term_pf2}
+	\Ebb(D_kf(x)) = \sum_{t = 1}^{s - 1} f_t(x) I_t + O(r^s) = \sum_{t = 2}^{s - 1} f_t(x) I_t + O(r^s),
+	\end{equation}
+	which establishes~\eqref{eqn:leading_term} in the base case.
+
+	Now, we assume \eqref{eqn:leading_term} holds for all $k \in [n]^q$, and prove the desired estimate on $\Ebb(D_{kj}f(x))$ for each $j \in [n]$.
+By the law of iterated expectation and~\eqref{eqn:leading_term_pf2},
+	\begin{align*}
+	\Ebb(D_{kj}f(x)) & = \Ebb(D_k(\Ebb(D_jf))(x)) \\
+	& = \Ebb\left(D_k\left(\sum_{t = 2}^{s - 1} I_t f_t + O(r^s)\right)(x)\right) \\
+	& = \sum_{t = 2}^{s - 1} I_t \cdot \Ebb(D_kf_t(x)) + O(r^s)
+	\end{align*}
+	where the second equality follows from the linearity and boundedness of $f \mapsto \Ebb(D_kf)$.
+We now apply the inductive hypothesis to $\Ebb(D_kf_t(x))$.
+If $2(q + 1) \geq s$, note that since $f_t \in C^{s - t}(L)$ for $t \geq 2$, we have by hypothesis $\Ebb(D_kf_t(x)) = O(r^{s - t})$.
+As a result
+	\begin{equation*}
+	\sum_{t = 2}^{s - 1} I_t \cdot \Ebb(D_kf_t(x)) = \sum_{t = 2}^{s - 1} I_t \cdot O(r^{s - t}) = O(r^s)
+	\end{equation*} 
+	Otherwise $2(q + 1) \leq s - 1$.
+For each $t = 2,\ldots, s - 1$, if additionally  $2q \leq s - t - 1$, then by hypothesis $\Ebb(D_kf_t(x)) = \sum_{\ell = 2q}^{s - t - 1} O(r^{\ell}) \cdot g_{\ell}(x) + O(r^{s - t})$ for some $g_{\ell} \in C^{s - t - \ell}(L)$, and otherwise $\Ebb(D_kf_t(x)) = O(r^{s - t})$.
+Therefore,
+	\begin{align*}
+	\sum_{t = 2}^{s - 1} I_t \cdot \Ebb(D_kf_t(x)) & = \sum_{t = 2}^{s - 1 - 2q} I_t \cdot \left\{\sum_{\ell = 2q}^{s - t - 1} O(r^{\ell}) \cdot g_{\ell}(x) + O(r^{s - t}) \right\} + \sum_{t = s - 1 - 2q}^{s - 1}I_{t} \cdot O(r^{s - t}) \\
+	& = \sum_{t = 2}^{s - 1 - 2q} I_t \cdot \left\{\sum_{\ell = 2q}^{s - t - 1} O(r^{\ell}) \cdot g_{\ell}(x)\right\}  + O(r^s) \\
+	& = \sum_{\ell = 2q}^{s - 3} \sum_{t = 2}^{s - \ell - 1} I_t \cdot O(r^{\ell}) \cdot g_{\ell}(x) + O(r^s).
+	\end{align*}
+	Noting that $g_{\ell} \in C^{s - (t + \ell)}(L)$ for some $\ell + t = 2(q + 1),\ldots,s - 1$, and $I_t \cdot O(r^{\ell}) = O(r^{t + \ell})$, we can rewrite the final equation as a sum over $\ell + t = 2(q + 1),\ldots,s - 1$, which proves~\eqref{eqn:leading_term}.
+\end{proof}
+
+\section{Old Work}
+
+\subsection{Bounding the 2nd-Order Roughness Functional}
+The (first and) second order roughness functionals are somewhat special.
+These roughness functionals do not involve taking compositions of the difference operator $D_k$, and as a result, we get the desired bounds in expectation even when $r(n)$ is small.
+\begin{lemma}
+	\label{lem:2nd_order_roughness_functional}
+	Fix $L > 0$, and form the neighborhood graph $G_{n,r}$ using kernel $K(z) = \1\{z^2 \leq r^2\}$. Suppose $f \in C^2(L)$ and $p \in C^1(L)$. Then for any $r(n) \geq n^{-1/(2+d)}$ and , we have
+	\begin{equation*}
+	\Ebb(R_{2,n}(f)) = O(L^2)
+	\end{equation*}
+\end{lemma}
+
+\subsection{Bounding the 3rd-Order Roughness Functional}
+
+For the order-$s$ roughness functionals when $s \geq 3$, the graph Laplacian is forced to approximate higher order derivatives by composing difference operators-- that is, taking differences of differences, etc.
+In for the roughness functional to continue scaling at the desired rate, we will need $r(n)$ to be larger, and we cannot use the uniform kernel $K$ any longer.
+In the particular case $s = 3$, we will require a kernel $K$ which is bounded and compactly supported on $B(0,1)$, and additionally satisfies
+\begin{equation*}
+\int z^j K(z) = 0~~\textrm{for $j = 1,2$.}
+\end{equation*}
+We then have the following result.
+\begin{lemma}
+	\label{lem:3rd_order_roughness_functional}
+	Suppose $f \in C^{3}(L)$ for some $L > 0$, and further suppose $p \in C^{2}(L)$. For any kernel $K$ which satisfies the previous conditions, and any $r(n) \geq n^{-1/(4+d)}$, we have that
+	\begin{equation*}
+	\Ebb(R_{s,n}(f)) = O(1)
+	\end{equation*} 
+\end{lemma}
 
@@
 
-\section{Proof of Lemma~\ref{lem:3rd_order_roughness_functional}}
+\subsection{Proof of Lemma~\ref{lem:3rd_order_roughness_functional}}
 We prove Lemma~\ref{lem:3rd_order_roughness_functional} when $d = 1$ and $r(n) \geq n^{-1/5}$.
@@
 The above four cases cover all terms in the summation, and so we have proved Lemma~\ref{lem:3rd_order_roughness_functional}.
-\section{Proof of Lemma~\ref{lem:expected_difference_operators}}
-
-\subsection{Proof of~\eqref{eqn:expected_difference_operators_1}}
-
-Let $k,\ell \in [n]^q$ be index vectors. We first prove the desired bound in the case when some indices are repeated, and then the desired bound in the case when all indices are distinct.
-
-\paragraph{Repeated indices.}
-
-Since $f \in C^1(L)$, for all index vectors $k,\ell \in [n]^q$ and indices $i \in [n]$, the product of iterated difference operators $D_kf(x_i) D_{\ell}f(x_i)$ satisfies
-\begin{equation*}
-\abs{D_kf(x_i) D_{\ell}f(x_i)} \leq 4^{q} L^2 r^{2 - 2q}
-\end{equation*}
-Moreover $D_kf(x_i) D_{\ell}f(x_i)$ will equal zero if there exists $x_j, j \in k \cup \ell \cup i$ such that
-\begin{equation*}
-\norm{x_j - x_h} > r,~~\textrm{for all $h \in k \cup \ell \cup i$}
-\end{equation*}
-Therefore $D_kf(x_i) D_{\ell}f(x_i)$ is nonzero with probability $O(r^{\abs{k \cup \ell \cup i} - 1})$, which along with the boundedness of $D_kf(x_i) D_{\ell}f(x_i)$ implies the claimed result.
-
-\paragraph{All indices distinct.}
-
-By the law of iterated expectation, we have
-\begin{align}
-\Ebb\left[D_kf(x_i)D_{\ell}f(x_i)\right] & = \Ebb\left[\Ebb\left(D_kf(x)|x_i = x\right) \Ebb\left(D_{\ell}f(x)|x_i = x\right)\right] \nonumber \\
-& = \Ebb\left[\Ebb\left(D_kf(x)|x_i = x\right)^2\right] \label{eqn:expected_difference_operators_pf1}
-\end{align}
-Again applying the law of iterated expectation, letting $g$ be the function given by $g(x) = \Ebb(D_1f(x))$, we have
-\begin{align*}
-\Ebb(D_kf(x)) & = \Ebb(\Ebb(D_{k}f(x)|x_{k_1},\ldots,x_{k_{q-1}})) \\
-& = \Ebb((D_{k_1\cdots k_{q - 1}}g)(x)).
-\end{align*}
-By Lemma~\ref{lem:leading_term}, we have that $g(x) = O(r^s)$ for all values of $x$, and therefore
-\begin{equation*}
-(D_{k_1\cdots k_{q_1}}g)(x) = O(r^{s - d(q - 1)})
-\end{equation*}
-Since $(D_{k_1\cdots k_{q_1}}g)(x)$ is nonzero with probability $O(r^{d(q - 1)})$, we have that 
-\begin{equation*}
-\Ebb(D_kf(x)) = \Ebb((D_{k_1\cdots k_{q - 1}}g)(x)) = O(r^s),
-\end{equation*}
-and the desired result follows from~\eqref{eqn:expected_difference_operators_pf1}.
-
-\subsection{Proof of~\eqref{eqn:expected_difference_operators_2}}
-
-The proof of~\eqref{eqn:expected_difference_operators_2} is essentially the same, and we do not reproduce it here.
-\section{Auxiliary Results.}
-\begin{lemma}
-	\label{lem:difference_operator_taylor}
-	When $f \in C^s(L)$, for any $k \in [n]$ 
-	\begin{equation*}
-	D_kf(x) = \left(\sum_{r = 1}^{s - 1} (x_k - x)^r f^{(r)}(x) + O((x_k - x)^{s})\right) \cdot K_r(x_k,x)
-	\end{equation*}
-\end{lemma}
-
-\begin{lemma}
-	\label{lem:expected_kernel}
-	When $K$ is a bounded and compactly supported kernel, then for any $s$ and for all $x \in \Rd$
-	\begin{equation*}
-	\Ebb(\abs{K(x_i,x)}^s) \leq CLr^{(1 - s)d}
-	\end{equation*}
-\end{lemma}
-
-\begin{lemma}
-	\label{lem:leading_term}
-	When $f \in C^s(L)$ and $p \in C^{s - 1}(L)$, we have that 
-	\begin{equation*}
-	\Ebb(\left(f(x_{\ell}) - f(x)\right)K_r(x_{\ell},x)) \leq Lr^s.
-	\end{equation*}
-\end{lemma}
 

### 2020-01-25 notes/graph_testing.tex
updated weekly notes and project level notes

--- 2019-12-20 notes/graph_testing.tex
@@
 
-In Theorem~\ref{thm:sobolev_testing_rate} we show that under some typical regularity conditions on $P$, the test $\phi_{\textrm{spec}} := \1\{T_{\mathrm{spec}} \geq \tau\}$ is, up to log factors, a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};R)$.
-To conveniently state our results we introduce the notation
-\begin{equation*}
-h(a,d) = a(d+2) + (1 + 2/d),~~\textrm{for $a > 0$}
-\end{equation*}
+In Theorem~\ref{thm:sobolev_testing_rate} we show that under some typical regularity conditions on $P$, the graph Laplacian eigenvector projection test $\phi_{\textrm{spec}} := \1\{T_{\mathrm{spec}} \geq \tau\}$ is a minimax optimal test over the Sobolev balls $W^{s,2}(\mathcal{X};R)$ whenever $4s > d$ and $\Xset = [0,1]^d$.
 
@@
 	\label{thm:sobolev_testing_rate}
-	Let $b \geq 1$ and $a > 0$ be fixed constants, and let $d < 4$.
-Suppose that $\Pbb$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p$ bounded above and below by constants, i.e
-	\begin{equation*}
-	0 < p_{\min} < p(x) < p_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
+	Let $b \geq 1$ and $R > 0$ be fixed constants.
+Suppose that $\Pbb$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p \in C^{s-1}(\Xset;p_{\max})$ for some $p_{\max} < \infty$, and further that $p(x)$ is bounded away from zero, i.e.
+there exists $p_{\min} > 0$ such that
+	\begin{equation*}
+	p_{\min} < p(x),~~ \textrm{for all $x \in \mathcal{X}$.}
 	\end{equation*}
@@
 	\begin{equation*}
-	r = \log^a n \cdot \left(\frac{\log n}{n}\right)^{1/d}, ~\kappa = n^{2d/(4 + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
-	\end{equation*}
-	then there exists a constant $c$ which may depend on $d,R,p_{\max},b$ and $a$ but is independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
+	\textrm{$K$ the uniform kernel,}~ n^{-1/(2(s-1) + d)} \leq r(n) \leq n^{-4/((4s + d)(2+d))}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	\end{equation*}
+	then there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
 	\begin{equation}
 	\label{eqn:sobolev_testing_rate}
-	\epsilon^2 \geq c \cdot b \cdot n^{-4/(4 + d)} (\log n)^{h(a,d)}
+	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4s + d)}\} \cdot n^{-{4s}/(4s + d)}
 	\end{equation}
@@
 	\label{eqn:sobolev_testing_rate_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; \mathcal{W}^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}.
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W^{2,s}(\mathcal{X};R)) \leq \frac{c_2}{b}.
 	\end{equation}
 \end{theorem}
-The testing rate~\eqref{eqn:sobolev_testing_rate} matches the minimax critical radius up to a factor of $\log^{h(a,d)}n$.
-
-When $d \geq 4$ the compact embedding
-\begin{equation*}
-W_d^{1,2}(\mathcal{X}) \subseteq \mathcal{L}_d^4(\mathcal{X}) 
-\end{equation*}
-does not hold (as it does when $d < 4$). However, if we directly assume $f \in \mathcal{L}_d^4(\mathcal{X};R)$ (regardless of $d$) we obtain the following result.
+
+The restriction that $p$ be supported on the unit cube is mostly for convenience.
+If instead $p$ is supported on any compact set with Lipschitz boundary, the Theorem statement should hold (up to constants), with only a few modifications to our proofs.
+However, we do not work through the details.
+
+The restriction $4s > d$ is a fundamental consequence of the tightness of the Sobolev embedding theorem. When $4s \geq d$ the compact embedding
+\begin{equation*}
+W_d^{s,2}(\mathcal{X}) \subseteq \mathcal{L}_d^4(\mathcal{X}) 
+\end{equation*}
+does not hold (as it does when $d < 4s$). However, if we directly assume $f \in \mathcal{L}_d^4(\mathcal{X};R)$ (regardless of $d$) we obtain the following result.
 \begin{proposition}
 	\label{prop:L4_testing_rate}
-	If $f \in \mathcal{L}_d^4(\mathcal{X};R)$, there exists a constant $c$ such that if
+	If $f \in \Leb^4(\mathcal{X};R)$, there exists a constant $c$ such that if
 	\begin{equation}
@@
 \end{proposition}
-
-Note that when $d < 4$ the rate~\eqref{eqn:sobolev_testing_rate} is sharper than \eqref{eqn:L4_testing_rate}.
+Note that when $4s > d$ the critical radius~\eqref{eqn:sobolev_testing_rate} is smaller than \eqref{eqn:L4_testing_rate}.
 
@@
 	test over the Sobolev ball $W_d^{1,2}(\mathcal{X};R)$?
-	\item Does the testing rate of $\phi_{\mathrm{spec}}$ over Sobolev spaces $W_d^{s,2}(\mathcal{X};R)$ match the minimax rate, for an appropriate choice of kernel $K$?
 	\item Assume the distribution $P$ is supported on a manifold $\mathcal{M}$ of intrinsic dimension $s < d$.
@@
 
-\section{Definitions}
+\section{Definitions and Notation}
 
@@
 
-\subsection{Graph Operators.}
+\subsection{Graph Operators}
 Let $s \geq 1$ be an integer. The $s$th-order difference operator on $G_{n,r}$, denoted $B^{(s)}$, is defined by
@@
 
-\section{\textcolor{red}{In Progress}}
-
-We say $K_{r}$ is an order-$s$ kernel when 
-\begin{equation*}
-\int x^j K_{r}(x) \,dx = 0, \textrm{for $j = 1,\dots,s$}
-\end{equation*}
-
-Our goal is to prove the following theorem.
-
-\begin{theorem}
-	\label{thm:higher_order_sobolev_testing_rate}
-	Let $b,s,d \geq 1$ and $a > 0$ be fixed constants such that $d < 4s$.
-Suppose that $\Pbb$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p$ bounded above and below by constants, i.e
-	\begin{equation*}
-	0 < p_{\min} < p(x) < p_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
-	\end{equation*}
-	Then the following statement holds: if the test $\phi_{\spec}$ is performed with parameter choices 
-	\begin{equation*}
-	\textrm{$K_r$ an order-$s$ kernel with}~ r = \log^a n \cdot \left(\frac{\log n}{n}\right)^{1/d}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
-	\end{equation*}
-	then there exists a constant $c$ which may depend on $d,R,p_{\max},b$ and $a$ but is independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
-	\begin{equation}
-	\label{eqn:higher_order_sobolev_testing_rate}
-	\epsilon^2 \geq c \cdot b \cdot n^{-4/(4s + d)} (\log n)^{h(a,d)}
-	\end{equation}
-	the worst-case risk is upper bounded
-	\begin{equation}
-	\label{eqn:higher_order_sobolev_testing_rate_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; \mathcal{W}^{s,2}(\mathcal{X};R)) \lesssim \frac{1}{b}.
-	\end{equation}
-\end{theorem}
-
-\section{\textcolor{red}{In Progress Proofs}}
-
-\subsection{Proof of Theorem~\ref{thm:higher_order_sobolev_testing_rate}}
-
-Let $G = (V,E)$ be a graph over vertices $V = \set{v_1,\ldots,v_n}$, and let $\beta = (\beta_1,\ldots,\beta_n) \in \Reals^n$ be a signal over the vertices $V$.
-We observe responses $Y = (y_1,\ldots,y_n)$ according to the model
-\begin{equation*}
-y_i = \beta_i + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
-\end{equation*}
-Letting $L = VSV^T$ be the spectral decomposition of the Laplacian $L$ of $G$, our graph spectral test statistic is 
-\begin{equation*}
-T_{\spec} = \frac{1}{n}\sum_{k = 1}^{\kappa} \left(\sum_{i = 1}^{n} v_i y_i\right)^2
-\end{equation*}
-where $\kappa$ is a tuning parameter. The resulting test we will use is
-\begin{equation*}
-\phi_{\spec} = \1\{T_{\spec} \geq \frac{\kappa}{n} + t(b)\},~~\textrm{where}~ t(b) = b\sqrt{\frac{2\kappa}{n^2}}~~\textrm{for $b \geq 1$.}
-\end{equation*}
-
-Let $S_s(\beta;G)$ be a measure of smoothness the signal $\beta$ displays over the graph $G$, given by
-\begin{equation*}
-S_s(\beta;G) := \beta^T L^s \beta
-\end{equation*}
-In Lemma~\ref{lem:higher_order_fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}$.
-Our bound on the Type II error will be stated as a function of $S_2(\beta;G)$ as well as the $\kappa$th eigenvalue $s_{\kappa}$.
-
-\begin{lemma}
-	\label{lem:higher_order_fixed_graph_testing}
-	Let $1 \leq \kappa \leq n$ be an integer.
-	\begin{enumerate}
-		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}$ is upper bounded
-		\begin{equation}
-		\label{eqn:higher_order_graph_spectral_type_I_error}
-		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{1}{b^2}.
-		\end{equation}
-		\item \textbf{Type II error:} For any $b$ and $\beta$ such that
-		\begin{equation}
-		\label{eqn:higher_order_fixed_graph_testing_critical_radius}
-		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_s(\beta;G)}{ns_{\kappa}^s}
-		\end{equation}
-		the Type II error of $\phi_{\spec}$ is upper bounded,
-		\begin{equation}
-		\label{eqn:higher_order_graph_spectral_type_II_error}
-		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{3}{b}.
-		\end{equation}
-	\end{enumerate}
-\end{lemma}
-
-To prove Lemma~\ref{lem:higher_order_fixed_graph_testing} we will first compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}$, and then use Chebyshev's inequality to show \eqref{eqn:higher_order_graph_spectral_type_I_error} and \eqref{eqn:higher_order_graph_spectral_type_II_error}.
-
-\vspace{.2 in}
-
-\textit{Mean of $T_{\mathrm{spec}}$:} Using the notation $\dotp{v}{w} = \sum_{i = 1}^{n} v_iw_i$, we have
-\begin{align*}
-\Ebb(T_{\spec}) & = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{\beta}{v_k}^2 + \Ebb\bigl( \dotp{\varepsilon}{v_k}^2 + 2 \dotp{\varepsilon}{v_k} \dotp{\beta}{v_k}\bigr)\right) \\
-& = \frac{\kappa}{n} + \frac{1}{n}\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2.
-\end{align*}
-When $\beta = 0$, this equals $\kappa/n$. Otherwise, we have the following lower bound:
-\begin{align*}
-\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2 & = \norm{\beta}_2^2 - \sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \\
-& \geq \norm{\beta}_2^2 - \frac{1}{s_{\kappa}^s}\sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 s_k^s \\
-& \geq \norm{\beta}_2^2 - \frac{S_s(\beta;G)}{s_{\kappa}^s},
-\end{align*}
-and therefore $\Ebb(T_{\spec}) \geq \kappa/n + n^{-1}(\norm{\beta}_2^2 - S_s(\beta;G)/s_{\kappa}^s)$.
-
-\vspace{.2 in}
-
-\textit{Variance of $T_{\mathrm{spec}}$:}
-We write $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvectors $v_1,\ldots,v_{\kappa}$ as columns. Consequently,
-\begin{align}
-\Var(T_{\spec}) & = \frac{1}{n^2} \Var(y^T V_{\kappa} V_{\kappa}^T y) \\
-& = \frac{1}{n^2} \Var((\beta + \varepsilon)^T V_{\kappa} V_{\kappa}^T (\beta + \varepsilon)) \\
-& = \frac{1}{n^2} \Var(2 \beta^T V_{\kappa} V_{\kappa}^T \varepsilon + \varepsilon^T V_{\kappa} V_{\kappa}^T \varepsilon) \\
-& \leq \frac{1}{n^2}(4 \beta^T V_{\kappa} V_{\kappa}^T \beta + 2\kappa)
-\end{align}
-where the last inequality follows from standard properties of the Gaussian distribution.
-We now move on to showing the desired inequalities \eqref{eqn:higher_order_graph_spectral_type_I_error} and \eqref{eqn:higher_order_graph_spectral_type_II_error}.
-
-\vspace{.2 in}
-
-\textit{Proof of~\eqref{eqn:higher_order_graph_spectral_type_I_error}:} By Chebyshev's inequality,
-\begin{align*}
-\Pbb_{\beta = 0}\bigl(T_{\spec} \geq \frac{\kappa}{n} + t(b)\bigr)
-& \leq \Pbb_{\beta = 0}\bigl(\abs{T_{\spec} - \frac{\kappa}{n}} \geq t(b)\bigr) \\
-& \leq \frac{\Var_{\beta = 0}(T_{\spec})}{t(b)^2} = \frac{1}{b^2}.
-\end{align*}
-
-\vspace{.2 in}
-
-\textit{Proof of~\eqref{eqn:higher_order_graph_spectral_type_II_error}:} For simplicity, we introduce the notation
-\begin{equation*}
-\Delta = \frac{\norm{\beta}_2^2}{n} - \frac{S_s(\beta;G)}{ns_{\kappa}^s}.
-\end{equation*}
-Assumption~\eqref{eqn:higher_order_fixed_graph_testing_critical_radius} implies $\Delta \geq 2 t(b)$. Then another application of Chebyshev's inequality gives us
-\begin{align*}
-\Pbb_{\beta}\bigl(T_{\spec} \leq \frac{\kappa}{n} + t(b)\bigr) & = \Pbb_{\beta}\bigl(T_{\spec} - \Ebb_{\beta}(T_{\spec}) \leq t(b) - \Delta \bigr) \\
-& \leq \Pbb_{\beta}\bigl(\abs{T_{\spec} - \Ebb_{\beta}(T_{\spec})} \leq \Delta - t(b) \bigr) \tag{since $\Delta \geq t(b)$}	\\
-& \leq \frac{\Var_{\beta}(T_{\spec})}{(\Delta - t(b))^2} \\
-& \leq 4\frac{\Var_{\beta}(T_{\spec})}{\Delta^2} \tag{since $\Delta \geq 2t(b)$} \\
-& \leq 4\frac{2\kappa/n^2 + 4\beta^T V_{\kappa} V_{\kappa}^T \beta /n^2}{\Delta^2}.
-\end{align*}
-We handle each summand in the numerator separately. For the first term, since $\Delta \geq 2 t(b)$, we have
-\begin{equation}
-\label{eqn:higher_order_spectral_type_II_error_pf1}
-\frac{2\kappa}{n^2\Delta^2} \leq \frac{1}{2b^2}.
-\end{equation}
-
-For the second term, noting that $\Delta = \beta^T V_{\kappa} V_{\kappa}^T \beta/n$, we have
-\begin{align}
-\frac{\beta^T V_{\kappa} V_{\kappa}^T \beta/n^2}{\Delta^2} & = \frac{1}{n\Delta} \nonumber \\
-& \leq \frac{1}{2nt(b)} \nonumber \\
-& = \frac{1}{2b\sqrt{2\kappa}}, \label{eqn:higher_order_spectral_type_II_error_pf2}
-\end{align}
-and combining~\eqref{eqn:higher_order_spectral_type_II_error_pf1} and~\eqref{eqn:higher_order_spectral_type_II_error_pf2} yields~\eqref{eqn:higher_order_graph_spectral_type_II_error}.
-
-\subsubsection{Step 2: Bounding neighborhood graph functionals}
-
-To make use of Lemma~\ref{lem:fixed_graph_testing} we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$.
-In particular, we will show that for some constants $c_1,c_2,c_3$, the following statements:
-\begin{enumerate}[label=(E\arabic*)]
-	\item 
-	\label{event:higher_order_discrete_sobolev_norm}
-	\textbf{Graph Sobolev norm:} When $\beta = (f(x_1),\ldots,f(x_n))$ for $f \in W^{s,2}(\mathcal{X};R)$, and $r = n^{-1/(2 + d)}$,
-	\begin{equation}
-	\label{eqn:higher_order_continuous_to_discrete_sobolev_norm}
-	S_s(\beta;G_{n,r}) \leq c_1 \cdot n^{2 + s} r^{s(d + 2)} 
-	\end{equation}
-	\item 
-	\label{event:higher_order_eigenvalue_tail_decay}
-	\textbf{Eigenvalue tail bound:} For any $r \geq n^{-1/(2 + d)}$ and for all $\kappa \in [n]$:
-	\begin{equation}
-	\label{eqn:higher_ordereigenvalue_tail_bound}
-	s_{\kappa} \geq c_2 \cdot n r^{d + 2} \kappa^{2/d}
-	\end{equation}
-	\item 
-	\label{event:higher_order_l2_norm}
-	\textbf{Empirical norm of $f$:} When $f \in W^{s,2}(\mathcal{X};R)$ and \textcolor{red}{some other condition}
-	\begin{equation}
-	\label{eqn:higher_order_l2_to_empirical_norm}
-	\frac{1}{n}\sum_{i = 1}^{n}\beta_i^2 = \norm{f}_n^2 \geq c_3 \cdot \norm{f}_{\Leb^2}^2
-	\end{equation}
-\end{enumerate}
-
-each hold with probability $1 - o(1)$ as $n \to \infty$.
+\subsection{Kernels}
+We say a kernel function $K(\cdot)$ is a $2$nd order kernel if $K$ is compactly supported on $B(0,1)$, uniformly upper bound $\abs{K(x)} \leq K_{\max} < \infty$ for all $x \in B(0,1)$, and
+\begin{equation*}
+\int K(x) \,dx = \nu_d,~~ \int x K(x) \,dx = 0.
+\end{equation*}
+Note that the uniform kernel, defined as $K(x) = \1\{\norm{x} \leq 1\}$, is a $2$nd order kernel.
+
+\subsection{Notation}
+
+We will treat $f$ interchangeably as a function $f:\Rd \to \Reals$, and as a vector $f = (f(x_1),\ldots,f(x_n))$. It should always be clear from context how we are using $f$.
+For sequences $(a_n), (b_n)$, we say $a_n = O(b_n)$ if there exists a constant $c$ such that $a_n \leq cb_n$ for all $c$. We use the notation $a \wedge b$ for the minimum of $a$ and $b$.
+
+We use the notation $\Lambda(H)$ to denote the spectrum of a matrix $H$, and $\lambda_k(H)$ to denote the $k$th smallest eigenvalue of $H$.
 
@@
 \subsection{Proof of Theorem~\ref{thm:sobolev_testing_rate}}
-
-To prove Theorem~\ref{thm:sobolev_testing_rate}, we will show that when
-\begin{equation}
-\label{eqn:critical_radius}
-\epsilon^2 \geq c \cdot b \cdot (\log n)^{h(a,d)} \cdot n^{-4/(4 + d)}
-\end{equation}
-and we choose tuning parameters as specified in Theorem~\ref{thm:sobolev_testing_rate}, the test $\phi_{\mathrm{spec}}$ satisfies
-\begin{equation}
-\mathcal{R}_{\epsilon}(\phi_{\spec}; W^{1,2}(\mathcal{X};R)) = \Ebb_{f_0}(\phi_{\spec}) + \sup_{\substack{f \in W^{1,2}(\mathcal{X};R) \\ \norm{f}_{\Leb^2(\mathcal{X})} > \epsilon}} \Ebb_{f}(1 - \phi_{\spec}) \leq \left(\frac{1}{b^2} + \frac{3}{b}\right) + o(1).
-\end{equation}
 
@@
 
-Let $G = (V,E)$ be a graph over vertices $V = \set{v_1,\ldots,v_n}$, and let $\beta = (\beta_1,\ldots,\beta_n) \in \Reals^n$ be a signal over the vertices $V$.
-We observe responses $Y = (y_1,\ldots,y_n)$ according to the model
-\begin{equation*}
-y_i = \beta_i + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
-\end{equation*}
-Letting $L = VSV^T$ be the spectral decomposition of the Laplacian $L$ of $G$, our graph spectral test statistic is 
+Let $G = (V,E)$ be a graph over $n$ vertices $V$, and let $\beta = (\beta_1,\ldots,\beta_n) \in \Reals^n$ be a signal over the vertices.
+(Although one can think of $G$ as a realization of $G_{n,r}$ for a particular sample $X$, our theory in this section will deal with $G$ as if it were an arbitrary fixed graph.) We observe responses $Y = (y_1,\ldots,y_n)$ according to the model
+\begin{equation*}
+y_i = \beta_i + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} N(0,1)
+\end{equation*}
+Let $L$ denote the graph Laplacian of $G$, and let $0 = \lambda_1 \leq \lambda_2 \leq \ldots \leq \lambda_n$ be the ordered eigenvalues of $L$, with eigenvector $v_i$ corresponding to eigenvalue $\lambda_i$.
+Our graph spectral test statistic is
 \begin{equation*}
@@
 
-Let $S_2(\beta;G)$ be a measure of smoothness the signal $\beta$ displays over the graph $G$, given by
-\begin{equation*}
-S_2(\beta;G) = \sum_{i = 1}^{n} \sum_{j = 1}^{n} (\beta_i - \beta_j)^2 \1((v_i,v_j) \in E).
-\end{equation*}
-In Lemma~, we upper bound the Type I and Type II error of the test $\phi_{\spec}$.
-Our bound on the Type II error will be stated as a function of $S_2(\beta;G)$ as well as the $\kappa$th eigenvalue $s_{\kappa}$.
+In Lemma~\ref{lem:fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}$.
+Our bound on the Type II error will be stated as a function of $\beta^T L^s \beta$--a measure of the smoothness the signal $\beta$ displays over the graph $G$--as well as the $\kappa$th eigenvalue $\lambda_{\kappa}$.
 
@@
 		\label{eqn:fixed_graph_testing_critical_radius}
-		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_2(\beta;G)}{ns_{\kappa}}
+		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{\beta^T L^s \beta}{n\lambda_{\kappa}^s}
 		\end{equation}
@@
 
-The first term on the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius} grows with $\kappa$, while the second term shrinks.
-These can be thought of as the variance and squared bias terms, respectively, of the error.
-
-\paragraph{Proof of Lemma~\ref{lem:fixed_graph_testing}.}
-
 To prove Lemma~\ref{lem:fixed_graph_testing} we will first compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}$, and then use Chebyshev's inequality to show \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.
@@
 \sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2 & = \norm{\beta}_2^2 - \sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \\
-& \geq \norm{\beta}_2^2 - \frac{1}{s_{\kappa}}\sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 s_k \\
-& \geq \norm{\beta}_2^2 - \frac{S_2(\beta;G)}{s_{\kappa}},
-\end{align*}
-and therefore $\Ebb(T_{\spec}) \geq \kappa/n + n^{-1}(\norm{\beta}_2^2 - S_2(\beta;G)/s_{\kappa})$.
+& \geq \norm{\beta}_2^2 - \frac{1}{\lambda_{\kappa}^s}\sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \lambda_k^s \\
+& \geq \norm{\beta}_2^2 - \frac{\beta^T L^s \beta}{\lambda_{\kappa}^s},
+\end{align*}
+and therefore $\Ebb(T_{\spec}) \geq \kappa/n + n^{-1}(\norm{\beta}_2^2 - \beta^T L^s \beta/\lambda_{\kappa}^s)$.
 
@@
 \textit{Variance of $T_{\mathrm{spec}}$:}
-We write $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvectors $v_1,\ldots,v_{\kappa}$ as columns. Consequently,
-\begin{align}
+We write $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvector $v_k$ as its $k$th column.
+\begin{align*}
 \Var(T_{\spec}) & = \frac{1}{n^2} \Var(y^T V_{\kappa} V_{\kappa}^T y) \\
@@
 & \leq \frac{1}{n^2}(4 \beta^T V_{\kappa} V_{\kappa}^T \beta + 2\kappa)
-\end{align}
+\end{align*}
 where the last inequality follows from standard properties of the Gaussian distribution.
@@
 \begin{equation*}
-\Delta = \frac{\norm{\beta}_2^2}{n} - \frac{S_2(\beta;G)}{ns_{\kappa}}.
+\Delta = \frac{\norm{\beta}_2^2}{n} - \frac{\beta^T L^s \beta}{n\lambda_{\kappa}^s}.
 \end{equation*}
@@
 To make use of Lemma~\ref{lem:fixed_graph_testing} we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$.
-In particular, we will show that for some constants $c_1,c_2,c_3$, the following statements:
+In particular, we will show that the following statements each hold with probability at least $1 - c/b$ for sufficiently large $n$ (Here and in what follows, $c$ denotes a constant which is fixed in $n$ and $f$ and does not depend on $b$, but may depend on other fixed quantities such as $d$, $s$, etc.):
 \begin{enumerate}[label=(E\arabic*)]
@@
 	\label{event:discrete_sobolev_norm}
-	\textbf{Graph Sobolev norm:} When $\beta = (f(x_1),\ldots,f(x_n))$ for $f \in W^{1,2}(\mathcal{X};R)$:
+	\textbf{Graph Sobolev norm:} For any $n^{-1/(2(s - 1) + d)}\leq r \leq 1$,
 	\begin{equation}
 	\label{eqn:continuous_to_discrete_sobolev_norm}
-	S_2(\beta;G_{n,r}) \leq c_1 \cdot n^2 r^{d + 2} 
+	f^T L^s f \leq c \cdot b \cdot \norm{f}_{W^{s,2}(\Xset)}^2 \cdot n^{s + 1} r^{s(d + 2)} 
 	\end{equation}
@@
 	\label{event:eigenvalue_tail_decay}
-	\textbf{Eigenvalue tail bound:} If $r = (\log^{a}n)\cdot(\log n/n)^{1/d}$ for any $a > 0$, then for all $\kappa \in [n]$:
+	\textbf{Eigenvalue tail bound:} For any $\kappa = 1,\ldots,n$ and $r \leq \kappa^{-2/(d(2 +d))}$, 
 	\begin{equation}
 	\label{eqn:eigenvalue_tail_bound}
-	s_{\kappa} \geq c_2 \cdot \frac{\kappa^{2/d}}{n^{2/d}}
-	\end{equation}
+	\lambda_{\kappa} \geq c \cdot n r^{d + 2} \kappa^{2/d}
+	\end{equation}
+	In particular, the inequality~\eqref{eqn:eigenvalue_tail_bound} is satisfied when $\kappa = n^{2d/(4s + d)}$ for any $r \leq n^{-4/((2+d)(4s + d))}$. 
 	\item 
 	\label{event:l2_norm}
-	\textbf{Empirical norm of $f$:} When $f \in W^{1,2}(\mathcal{X};R)$ and one of (a) $d = 1,2$, or (b) $d = 3$ and $\norm{f}_{\Leb^2}^2 \geq n^{-2/(d - 2)}$, then
+	\textbf{Empirical norm of $f$:} There exists a constant $c_1$ such that if $\norm{f}_{\Leb^2(\Xset)} \geq c_1 \cdot b \cdot n^{-ws/(4s + d)}$,
 	\begin{equation}
 	\label{eqn:l2_to_empirical_norm}
-	\frac{1}{n}\sum_{i = 1}^{n}\beta_i^2 = \norm{f}_n^2 \geq c_3 \cdot \norm{f}_{\Leb^2}^2
+	\norm{f}_n^2 \geq \frac{1}{b} \cdot \norm{f}_{\Leb^2}^2
 	\end{equation}
@@
 
-each hold with probability $1 - o(1)$ as $n \to \infty$.
-
 \paragraph{Proof of~\eqref{eqn:continuous_to_discrete_sobolev_norm}:}
 
-Observe that
-\begin{align*}
-\frac{1}{n^2}\Ebb(\beta^T L \beta) = \Ebb\left(\frac{1}{n^2} \sum_{i,j = 1}^{n} (\beta_i - \beta_j)^2 A_{ij}\right) & = \frac{(n - 1)}{n} \int_{\Xset} \int_{\Xset} (f(x) - f(y))^2K_r(x,y) \,dP(y) \,dP(x) \\
-& \leq p_{\max}^2 \frac{(n - 1)}{n} \int_{\Xset} \int_{\X} (f(x) - f(y))^2K_r(x,y) \,dy \,dx.
-\end{align*}
-We will show that for any $f \in \mathcal{W}^{1,2}(\X;R)$
-\begin{equation}
-\label{eqn:nonlocal_continuous_sobolev_norm}
-\int_{\Xset} \int_{\Xset} (f(x) - f(y))^2K_r(x,y) \,dy \,dx \leq c r^{d + 2} \norm{f}_{W^{1,2}(\mathcal{X})}
-\end{equation}
-whence the desired result~\eqref{eqn:continuous_to_discrete_sobolev_norm} follows by Markov's inequality and the assumption $f \in W^{1,2}(\mathcal{X};R)$.
-
-We begin by dealing with complications due to smoothing at the boundary of $\mathcal{X} = [0,1]^d$.
-Let $U$ be any bounded open set such that $\Xset \subset U$ is compactly contained in $U$.
-Note that as $\partial \Xset$ is $C^1$, by Theorem \ref{thm:evans_extension} there exists $g \in W^{1,2}(\Reals^d)$ such that
-\begin{enumerate}[(i)]
-	\item
-	\label{eqn:evans_extension_1}
-	$g = f$, $P$-almost-everywhere in $\Xset$
-	\item 
-	$g$ has support within $V$, and  
-	\item 
-	\label{eqn:sobolev_1_bound_2}
-	$\norm{g}_{W^{1,2}(\Rd)} \leq c \norm{f}_{W^{1,2}(\Xset)}$ for a constant $c$ which depends only on $\Xset$.
-\end{enumerate}
-As a result of \ref{eqn:evans_extension_1},
-\begin{equation}
-\label{eqn:sobolev_1_bound_1}
-\int_{\Xset} \int_{\Xset} (f(x) - f(y))^2K_r(x,y) \,dy \,dx \leq \int_{\Rd} \int_{\Rd} (g(x) - g(y))^2K_r(x,y) \,dy \,dx.
-\end{equation}
-
-Next we smooth $g$, so that we may work with ordinary partial derivatives.
-We let $\eta \in C^{\infty}(\Rd)$ be given by
-\begin{equation*}
-\eta(x) :=
-\begin{cases}
-C \exp \left\{\frac{1}{\norm{x}^2 - 1}\right\}& \quad \textrm{if $\norm{x}_2 \leq 1$} \\
-0 & \quad \textrm{if $\norm{x}_2 \geq 1$}
-\end{cases}
-\end{equation*}
-where the normalizing constant $C > 0$ is chosen so that $\int_{\Rd} \eta dx = 1$. Let $\eta_r(x) := (1/r^d) \eta(x/r)$. Then, the mollification of $g$ by $\eta_r$ is given by
-\begin{align*}
-g^r & := g \ast \eta_r \\
-& = \int_{\Rd} \eta_r(x - y) g(y) dy
-\end{align*}
-(Refer to \textcolor{red}{Evans}, Appendix C, Theorem 7 for a proof that $g^r \in C^{\infty}(\Rd)$.)
-Adding and subtracting within \eqref{eqn:sobolev_1_bound_1}, we have
-\begin{align}
-\int_{\Rd} & \int_{\Rd} (g(x) - g(y))^2K_r(x,y) \,dy \,dx \\
-& \leq 3 \int_{\Rd} \int_{\Rd} \bigl((g(x) - g^r(x))^2 + (g^r(x) - g^r(y))^2 + (g^r(y) - g(y))^2\bigr)K_r(x,y) \,dy \,dx \nonumber \\
-& = 6 \int_{\Rd} \int_{\Rd} \bigl((g(y) - g^r(y))^2 K_r(x,y) \,dy \,dx + 2 \int_{\Rd} \int_{\Rd} \bigl((g^r(x) - g^r(y))^2 K_r(x,y) \,dy \,dx \label{eqn:sobolev_1_bound_5}
-\end{align}
-We deal with each summand individually, beginning with the first one. Let $\nu_d = \pi^{d/2}/\Gamma(d/2 + 1)$ denote the volume of a ball with unit radius in $\Rd$. Then, we have
-\begin{align}
-\int_{\Rd} \int_{\Rd} \bigl((g(y) - g^r(y))^2 K_r(x,y) \,dy \,dx & = \int_{\Rd} \int_{B(x,r)} \bigl((g(y) - g^r(y))^2\,dy \,dx \nonumber \\
-& = \int_{\Rd} \int_{B(y,r)} \bigl((g(y) - g^r(y))^2\,dx \,dy \tag{by Tonelli's Theorem} \nonumber \\
-& = \nu_d r^d \int_{\Rd}\bigl((g(y) - g^r(y))^2 \,dy \\
-& \leq \nu_d r^{d + 2} \int_{\Rd} \norm{\nabla g(y)}^2 \,dy \label{eqn:sobolev_1_bound_4}
-\end{align}
-where the last line follows from Lemma \ref{lem:poincare_mollify}.
-
-We now turn our attention to the second summand. Note that as $g^r \in C^{\infty}(\Rd)$, we may apply Theorem \ref{thm:taylor_expansion} and obtain
-\begin{align*}
-(g^r(x) - g^r(y))^2 & = \left(\int_{0}^{1} \nabla g^r(x + t(y - x)) \cdot (y - x) \,dt\right)^2 \\
-& \leq \int_{0}^{1} \bigl(\nabla g^r(x + t(y - x)) \cdot (y - x)\bigr)^2 \,dt  \tag{Jensen's inequality} \\
-& \leq \norm{y - x}^2 \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 \,dt. \tag{Cauchy-Schwarz inequality}
-\end{align*}
-As a result, we have
-\begin{align*}
-\int_{\Rd} \int_{\Rd} \bigl((g^r(x) - g^r(y))^2 K_r(x,y) \,dy) \,dx & \leq \int_{\Rd} \int_{\Rd} \norm{y - x}^2  \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 K_r(x,y) \,dt \,dy \,dx \\
-& \leq r^{2} \int_{\Rd} \int_{\Rd} \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 K_r(x,y) \,dt \,dy \,dx \\
-& = r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + t(y - x))}^2 K_r(x,y) \,dx \,dt \,dy \\
-& = r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + tz)}^2 K_r(z) \,dx \,dt \,dz \tag{$z = y - x$}
-\end{align*}
-
-where we write $K_r(z) = \mathbf{1}(\norm{z} \leq r)$ in an abuse of notation. Next, we note that
-\begin{equation*}
-\int_{\Rd} \norm{\nabla g^r(x + tz)}^2 \,dx = \int_{\Rd} \norm{\nabla g^r(x)}^2 \,dx \leq \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
-\end{equation*}
-with the inequality following from Lemma \ref{lem:gradient_mollify_commute}. Therefore,
-\begin{align}
-r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + tz)}^2 K_r(z) \,dx \,dt \,dz & \leq r^{2} \int_{\Rd} K_r(z) \int_{0}^{1} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx  \,dt \,dz \nonumber \\
-& = \nu_d r^{2 + d} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx. \label{eqn:sobolev_1_bound_3}
-\end{align}
-By \eqref{eqn:sobolev_1_bound_5}, \eqref{eqn:sobolev_1_bound_4} and \eqref{eqn:sobolev_1_bound_3}, we have that 
-\begin{equation*}
-\int_{\Rd} \int_{\Rd} (g(x) - g(y))^2 K_r(x,y) \,dx \,dy \leq 9 \nu_d r^{d + 2} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
-\end{equation*}
-Then by \ref{eqn:sobolev_1_bound_2}, $\int_{\Rd} \norm{\nabla g(x)}^2 \,dx \leq c \int_{\Xset} \norm{\nabla f(x)}^2 \,dx$, and the desired result of \eqref{eqn:sobolev_1_bound_1} follows.
+The probabilistic bound~\eqref{eqn:continuous_to_discrete_sobolev_norm} follows from the more general Lemma~\ref{lem:roughness_functional_expectation_sobolev} by Markov's inequality. 
+\begin{lemma}
+	\label{lem:roughness_functional_expectation_sobolev}
+	Let $\Xset$ be a Lipschitz domain, and let $s \geq 1$ be an integer.
+Suppose that $f \in W^{s,2}(\Xset)$, and further that $p \in C^{s-1}(\Xset;p_{\max})$ for some constant $p_{\max}$.
+Then for any $2$nd-order kernel $K$ and any $n^{-1/(2(s - 1) + d)} \leq r(n) \leq 1$, the expected graph Sobolev seminorm is upper bounded
+	\begin{equation}
+	\label{eqn:roughness_functional_expectation_sobolev}
+	\Ebb\bigl[f^T L^s f\bigr] \leq c \cdot \norm{f}_{W_d^{s,2}(\Xset)}^2 \cdot n^{s + 1}r^{s(d + 2)}
+	\end{equation}
+\end{lemma}
+
+The proof of Lemma~\ref{lem:roughness_functional_expectation_sobolev} is lengthy, and we defer it to Section~\ref{sec:technical_lemma_proofs}.
 
@@
 
-For ease of notation we write $t = n^{1/d}$, and assume without loss of generality that $t$ is whole number. Let $\overline{G} = (\overline{X},\overline{E})$ be the lattice over $[0,1]^d$, formally
-\begin{align*}
-\overline{X} & = \bigl\{\overline{x}_k := \frac{1}{t}(k_1,\ldots,k_d): k \in [t]^d\bigr\} \\
-\overline{E} & = \bigl\{(\overline{x}_k, \overline{x}_{\ell}): \norm{k - \ell}_1 = 1\bigr\}.
-\end{align*}
-
-By Theorem 1.1 of \textcolor{red}{Garcia Trillos and Slepcev}, with probability at least $1 - n^{-1}$ there exists a bijection $\pi: \overline{V} \to X$ such that
-\begin{equation}
-\label{eqn:transport_distance}
-\max_{k \in [t]^d} \abs{\overline{x}_k - \pi(\overline{x}_k)} \leq c \left(\frac{\log n}{n}\right)^{1/d}
-\end{equation}
-Since our choice $r = \log^a(n) \cdot (\log n/n)^{1/d}$ implies $r = \omega(\left(\frac{\log n}{n}\right)^{1/d})$, for $n$ sufficiently large with probability $1 - n^{-1} = 1 - o(1)$ we have
-\begin{equation}
-\label{eqn:eigenvalue_tail_bound_pf1}
-(\overline{x}_k,\overline{x}_{\ell}) \in \overline{E} \Longrightarrow (\pi(\overline{x}_k),\pi(\overline{x}_{\ell})) \in E.
-\end{equation}
-Let $\overline{L}$ be the Laplacian of $\overline{G}$, with eigenvalues $\overline{s}_1 < \cdots \leq \overline{s}_n$. If \eqref{eqn:eigenvalue_tail_bound_pf1} holds, then for all $x \in \Reals^n$,
-\begin{equation*}
-x^T \overline{L} x \leq x^T L x \Longrightarrow \overline{s}_j \leq s_j~~ \textrm{for all}~ j \in [n].
-\end{equation*}
-The eigenvalues of the $d$-dimensional lattice are explicitly known.
-In particular when $j = h^d$ for some whole number $h$, we have $s_j \geq \lambda_j = 4 d \sin^2(\pi h/2t) \geq \pi^2 h^2/ (4 t^2)$, with a similar inequality (for possible different constants) holding when $j^{1/d}$ is not integral; this shows~\eqref{eqn:eigenvalue_tail_bound}.
+We prove~\eqref{eqn:eigenvalue_tail_bound} by comparing $G_{n,r}$ to the tensor product of a $d$-dimensional lattice and a complete graph.
+The latter is a highly structured graph with known eigenvalues, which as we will see are sufficiently lower bounded for our purposes.
+
+Let $\wt{r} = r/(3(\sqrt{d} + 1)), M = (1/\wt{r})^d, N = n\wt{r}^d$. Assume without loss of generality that $M$ and $N$ are integers. Additionally, for $t = n^{1/d}$ and $m = M^{1/d}$ let 
+\begin{equation*}
+\overline{X} = \set{\frac{1}{t}(k_1,\ldots,k_d): k \in [t]^d},~~ \overline{Z} = \set{\frac{1}{m}(j_1,\ldots,j_d): j \in [m]^d}.
+\end{equation*}
+For a given $\overline{z}_j \in \overline{Z}$, we write $Q(z_j) = m^{-1}[j_1 - 1,j_1] \times \cdots \times m^{-1}[j_d - 1,j_d]$ for the cube of side length $1/m$ with $z_j$ at one corner.
+
+Consider the graph $H = (\overline{X}, E_H)$, where $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$ if
+\begin{equation*}
+\textrm{there exists}~\ol{z}_i, \ol{z}_j \in \ol{Z}~\textrm{such that}~\ol{x}_k \in Q(\ol{z}_i),~ \ol{x}_\ell \in Q(\ol{z}_j),~\textrm{and}~\norm{i - j}_1 \leq 1.
+\end{equation*}
+On the one hand $H \cong \ol{G}^M_d \otimes K_N$ where $\ol{G}^M_d$ is the $d$-dimensional lattice on $M$ nodes, and $K_N$ is the complete graph on $N$ nodes.
+On the other hand, we now show that with high probability $G_{n,r} \succeq H$.
+If $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$, then there exist $\ol{z}_i, \ol{z}_j$ such that
+\begin{equation*}
+\norm{\ol{x}_k - \ol{x}_{\ell}}_2 \leq m^{-1} + \norm{\ol{x}_k - \ol{z}_{i}}_2 + \norm{\ol{x}_{\ell} - \ol{z}_{j}}_2 \leq \wt{r}(1 + \sqrt{d}) = r/3.
+\end{equation*}
+
+Assuming~\eqref{eqn:slepcev_transport_distance} holds, if $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$, then for sufficiently large $n$
+\begin{equation*}
+\norm{\pi(\ol{x}_k) - \pi(\ol{x}_\ell)}_2 \leq 2 c \left(\frac{\log n}{n}\right)^{1/d} + \frac{r}{3} \leq r,
+\end{equation*}
+implying that $(\pi(\ol{x}_k), \pi(\ol{x}_{\ell})) \in E$. Therefore, $G_{n,r} \succeq \ol{G}^M_d \otimes K_N$ whenever~\eqref{eqn:slepcev_transport_distance} holds.
+
+The eigenvalues of lattices and complete graphs are known to satisfy, respectively
+\begin{equation*}
+\lambda_k(\ol{G}^{M}_d) \geq \frac{k^{2/d}}{M^{2/d}}~\textrm{for $k = 0,\ldots,M - 1$},~~ \textrm{and}~\lambda_{j}(K_N) \geq N\1\{j > 0\}~\textrm{for $j = 0,\ldots,N-1$.}
+\end{equation*}
+and by standard facts regarding the eigenvalues of tensor product graphs, we have that the spectrum $\Lambda(H)$ satisfies
+\begin{equation*}
+\Lambda(H) = \set{N\lambda_k(\ol{G}^{M}_d) + M\lambda_j(K_N): \textrm{for $k = 0,\ldots,M - 1$ and $j = 0,\ldots,N-1$}}
+\end{equation*}
+For all $j = 1,\ldots,N-1$, we have that $M\lambda_j(K_N) = MN = n$. Therefore,
+\begin{align*}
+\lambda_{\kappa}(H) & \geq \{n \wedge N\lambda_{\kappa}(\ol{G}^{M}_d)\} \\
+& \geq \{n \wedge n\wt{r}^d\frac{\kappa^{2/d}}{M^{2/d}}\} \\
+& \geq \{n \wedge (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d}\} \\
+& \geq (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d},
+\end{align*}
+where the last inequality is satisfies since $r \leq \kappa^{-2/(d(d + 2))}$.
+
+Finally, Theorem~\ref{thm:slepcev_transport_distance} establishes that \eqref{eqn:slepcev_transport_distance} occurs with probability at least $1 - c/n$.
+Since $\lambda_{\kappa}(G_{n,r}) \geq \lambda_{\kappa}(H)$ when~\eqref{eqn:slepcev_transport_distance} is satisfied, this completes the proof of~\eqref{eqn:eigenvalue_tail_bound}.
 
@@
 
-To prove~\eqref{eqn:l2_to_empirical_norm} we will upper bound
-\begin{equation*}
-\mathbb{E}\biggl[\Bigl(\frac{1}{n}\sum_{i = 1}^{n}f^2(x_i)\Bigr)^2\biggr],
-\end{equation*}
-and then apply the Paley-Zygmund inequality (Lemma~\ref{lem:paley_zygmund}). We have
-\begin{align*}
-\mathbb{E}\biggl[\Bigl(\frac{1}{n}\sum_{i = 1}^{n}f^2(x_i)\Bigr)^2\biggr] & = \frac{(n-1)}{n}\mathbb{E}(f^2(x_1)) + \frac{\mathbb{E}[f^4(x_1)]}{n} \\
-& \leq p_{\max}^2 \left(\norm{f}_{\Leb^2}^2 + \frac{\norm{f}_{\Leb^4}^4}{n}\right) \\
-& \overset{(i)}{\leq} p_{\max}^2 \left(\norm{f}_{\Leb^2}^2 + c\frac{\norm{f}_{\Leb^2}^{(4 - d)}}{n}\right) \\
-& \overset{(ii)}{\leq}  p_{\max}^2 \left(\norm{f}_{\Leb^2}^2 + c\norm{f}_{\Leb^2}^2\right).
-\end{align*}
-Here $(i)$ follows from two facts, the first being the embedding $W^{1,2}(\mathcal{X};R) \subseteq \Leb^q(\mathcal{X};R')$ for $q = 2d/(d - 2)$ when $d > 2$ and $q = \infty$ when $d \leq 2$ and some constant $R'$, and the second being a standard result on interpolation of $\Leb^p$ spaces, namely that since $f \in \Leb^2(\mathcal{X}) \cap \Leb^q(\mathcal{X})$, then
-\begin{equation*}
-\norm{f}_{\Leb^4} \leq \norm{f}_{\Leb^2}^{\lambda} \cdot \norm{f}_{\Leb^q}^{1 - \lambda},~~\textrm{where}~ \frac{1}{4} = \frac{\lambda}{2} + \frac{1 - \lambda}{q}.
-\end{equation*} 
-Then $(ii)$ follows from our assumptions on $\norm{f}_{\Leb^2}^2$: either 
-\begin{enumerate}[(a)]
-	\item $d \leq 2$ and $\norm{f}_{\Leb^2}^2 \geq 1/n$ so $\norm{f}_{\Leb^2}^{(4 - d)} \leq \max(R^{4 - d}/n, \norm{f}_{\Leb^2}^2) \leq c\norm{f}_{\Leb^2}^2$, or
-	\item $d = 3$ and $\norm{f}_{\Leb^2}^2 \geq n^{-2/(d - 2)}$.
-\end{enumerate} 
-Applying Lemma~\ref{lem:paley_zygmund} to  $\norm{f}_n^2$ yields~\eqref{eqn:l2_to_empirical_norm}.
-
-\subsubsection{Step 3: Conclusion}
+\begin{lemma}
+	\label{lem:empirical_norm_sobolev}
+	Let $\Xset$ be a Lipschitz domain over which the density is upper and lower bounded 
+	\begin{equation*}
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty ~~\textrm{for all $x \in \Xset$,}
+	\end{equation*}
+	and let $f \in W^{s,2}(\Xset)$.Then for any $b \geq 1$, there exists $c_1$ such that if 
+	\begin{equation}
+	\norm{f}_{\Leb^2(\Xset)} \geq 
+	\begin{cases*}
+	c_1 \cdot b \cdot \norm{f}_{W^{s,2}(\Xset)} \cdot \max\{n^{-1/2},n^{-s/d}\},~~\textrm{if}~2s \neq d \\
+	c_1 \cdot b \cdot \norm{f}_{W^{s,2}(\Xset)} \cdot n^{-a/2},~~\textrm{if}~ 2s = d ~\textrm{for any}~ 0 < a < 1
+	\end{cases*}
+	\end{equation}
+	then,
+	\begin{equation}
+	\label{eqn:l2_to_empirical_norm_sobolev}
+	\mathbb{P}\left[\norm{f}_n^2 \geq \frac{1}{b} \Ebb\bigl[\norm{f}_n^2\bigr]\right] \geq 1 - \frac{5}{b}
+	\end{equation}
+	where $c_1$ and $c_2$ are constants which may depend only on $s$, $\Xset$, $d$, $p_{\min}$ and $p_{\max}$.
+\end{lemma}
+The lower bound~\eqref{eqn:l2_to_empirical_norm} results from the more general Lemma~\ref{lem:empirical_norm_sobolev}, which can be verified by checking the various orderings of $2s/(4s + d)$, $s/d$ and $1/2$ whenever $4s < d$.
+\begin{proof}[Proof of Lemma~\ref{lem:empirical_norm_sobolev}]
+	To prove~\eqref{eqn:l2_to_empirical_norm_sobolev} we will show
+	\begin{equation*}
+	\mathbb{E}\bigl[\norm{f}_n^4\bigr] \leq \left(1 + \frac{1}{b^2}\right)\cdot \left(\mathbb{E}\bigl[\norm{f}_n^2\bigr]\right)^2
+	\end{equation*}
+	whence the claim follows from the Paley-Zygmund inequality (Lemma~\ref{lem:paley_zygmund}).
+Since $p \leq p_{\max}$ is uniformly bounded, we can relate $\mathbb{E}\bigl[\norm{f}_n^4\bigr]$ to the $\Leb^4$ norm,
+	\begin{equation*}
+	\mathbb{E}\bigl[\norm{f}_n^4\bigr] = \frac{(n-1)}{n}\left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + \frac{\mathbb{E}\Bigl[\bigl(f(x_1)\bigr)^4\Bigr]}{n} \leq \left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + p_{\max}^2\frac{\norm{f}_{\Leb^4}^4}{n}.
+	\end{equation*}
+	We will use a Sobolev inequality to relate $\norm{f}_{\Leb^4}$ to $\norm{f}_{W_d^{s,2}(\Xset)}$.
+The nature of this inequality depends on the relationship between $s$ and $d$ (see Theorem 6 in Section 5.6.3 of \textcolor{red}{Evans} for a formal statement), so from this point on we divide our analysis into three cases: (i) the case where $2s > d$, (ii) the case where $2s < d$, and (iii) the borderline case $2s = d$.
+
+	\paragraph{Case 1: $2s > d$.}
+	When $2s > d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in C^{\gamma}(\overline{\Xset})$ for some $\gamma > 0$ which depends on $s$ and $d$, with the accompanying estimate
+	\begin{equation*}
+	\sup_{x \in \Xset} \abs{f(x)} \leq \norm{f}_{C^{\gamma}(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}.
+	\end{equation*}
+	Therefore,
+	\begin{align*}
+	\norm{f}_{\Leb^4}^4 & = \int_{\Xset} [f(x)]^4 \,dx \\
+	& \leq \left(\sup_{x \in \Xset} \abs{f(x)}\right)^2 \cdot \int_{\Xset} [f(x)]^2 \,dx \\
+	& \leq c \norm{f}_{W^{s,2}(\Xset)}^2 \cdot \norm{f}_{\Leb^2(\Xset)}^2.
+	\end{align*}
+	Since by assumption
+	\begin{equation*}
+	\norm{f}_{\Leb^2(\Xset)}^2 \geq c_1^2 \cdot b^2 \cdot \norm{f}_{W_d^{s,2}(\Xset)}^2 \cdot \frac{1}{n},
+	\end{equation*}
+	we have
+	\begin{equation*}
+	p_{\max}^2\frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c \norm{f}_{W^{s,2}(\Xset)}^2 \cdot \frac{\norm{f}_{\Leb^2(\Xset)}^4}{n \norm{f}_{\Leb^2(\Xset)}^2} \leq c \frac{\norm{f}_{\Leb^2(\Xset)}^4}{c_1^2 b^2} \leq \frac{\Ebb\bigl[\norm{f}_n^2\bigr]}{b^2},
+	\end{equation*}
+	where the last inequality follows by taking $c_1$ sufficiently large.
+
+	\paragraph{Case 2: $2s < d$.}
+	When $2s < d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in \Leb^q(\Xset)$ for $q = 2d/(d - 2s)$, and moreover that
+	\begin{equation*}
+	\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}.
+	\end{equation*}
+	Since $4 = 2\theta + (1 - \theta)q$ for $\theta = 2 - d/(2s)$, Lyapunov's inequality implies
+	\begin{equation*}
+	\norm{f}_{\Leb^4(\Xset)}^4 \leq \norm{f}_{\Leb^2}^{2\theta} \cdot \norm{f}_{\Leb^q(\Xset)}^{(1 - \theta)q} \leq c \norm{f}_{\Leb^2(\Xset)}^{4} \cdot \left(\frac{\norm{f}_{W^{s,2}(\Xset)}}{\norm{f}_{\Leb^2(\Xset)}}\right)^{d/s}.
+	\end{equation*}
+	By assumption, $\norm{f}_{\Leb^2(\Xset)} \geq c_1 b \norm{f}_{W^{s,2}(\Xset)} n^{-s/d}$, and therefore
+	\begin{equation*}
+	p_{\max}^2 \frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c\norm{f}_{\Leb^2(\Xset)}^4 \left(\frac{\norm{f}_{W^{s,2}(\Xset)}}{n^{s/d}\norm{f}_{\Leb^2(\Xset)}}\right)^{d/s} \leq \frac{c\norm{f}_{\Leb^2(\Xset)}^4}{c_1b^{d/s}} \leq \frac{\norm{f}_{\Leb^2(\Xset)}^4}{b^2}.
+	\end{equation*}
+	where the last inequality follows when $c_1$ is sufficiently large, and keeping in mind that $d/s > 2$ and $b \geq 1$.
+
+	\paragraph{Case 3: $2s = d$.}
+	Assume $f$ satisfies~\eqref{eqn:paley_zygmund_1} for a given $0 < a < 1$.
+When $2s = d$, since $\Xset$ is a Lipschitz domain we have that $f \in L^q(\Xset)$ for any $q < \infty$, with the accompanying estimate
+	\begin{equation*}
+	\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}.
+	\end{equation*}
+	In particular the above holds for $q = 2/(1 - a)$ when $1/2 < a < 1$, and for any $q > 4$ when $0 < a < 1/2$. Using Lyapunov's inequality as in the previous case then implies the desired result.
+\end{proof}
+
+\subsubsection{Step 3: Putting the pieces together}
 
@@
 
-Now, we bound Type II error under the assumption $f \in W^{1,2}(\mathcal{X};R)$ and 
+Now, we bound Type II error under the assumption $f \in W^{s,2}(\mathcal{X};R), p \in C^{s - 1}(\Xset;p_{\max})$ uniformly bounded away from $0$ and $\infty$ over $\Xset$, and 
 \begin{equation}
 \label{eqn:critical_radius_1}
-\norm{f}_{\Leb^2}^2 \geq \epsilon^2 = c \cdot b \cdot (\log n)^{h(\beta,d)} \cdot n^{-4/(4 + d)}.
-\end{equation}
-We verify that the conditions required to apply Step 2 are satisfied.
-In particular, we choose $r = (\log^a n) \cdot (\log n/n)^{1/d}$, and observe that~\eqref{eqn:critical_radius_1} implies $\norm{f}_{\Leb^2}^2 \geq n^{-1}$ when $d \leq 2$ and $\norm{f}_{\Leb^2}^2 \geq n^{-2/(d - 2)}$ when $2 < d < 4$.
-We may therefore apply our conclusions in Step 2; namely, that for every possible choice of $f$ there exists a good set $\mathcal{E}_f \subseteq \Xset^n$ with $\Pbb(\mathcal{E}_f) \geq 1 - o(1)$ such that each of \eqref{eqn:continuous_to_discrete_sobolev_norm}, \eqref{eqn:eigenvalue_tail_bound}, and \eqref{eqn:l2_to_empirical_norm} hold for all $X \subseteq \mathcal{E}_f$.
-Choosing $\kappa = n^{2d/(4 + d)}$ to balance the squared bias and variance terms on the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius}, we have that for all $X \subseteq \mathcal{E}_f$
-\begin{align*}
-2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_2(\beta;G)}{ns_{\kappa}} & \leq 2bn^{-4/(4+d)} + \frac{n r^{d + 2}}{s_{\kappa}} \tag{by \eqref{eqn:continuous_to_discrete_sobolev_norm}} \\
-& \leq 2bn^{-4/(4+d)} + c\cdot\frac{n r^{d + 2}}{n^{4/(4+d)}} \tag{by \eqref{eqn:eigenvalue_tail_bound}} \\
-& \leq 2bn^{-4/(4+d)} + c\cdot(\log^{a(d + 2) + (1 + 2/d)}n)\cdot n^{-4/(4+d)} \\
-& \leq \frac{1}{10}\norm{f}_{\Leb^2} \tag{for a suitably large choice of $c$ in \eqref{eqn:critical_radius_1}}\\
-& \leq \frac{1}{n}\sum_{i = 1} \beta_i^2. \tag{by \eqref{eqn:l2_to_empirical_norm}}
-\end{align*}
+\norm{f}_{\Leb^2}^2 \geq \epsilon^2 = c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4s + d)}\} \cdot n^{-4s/(4s + d)}.
+\end{equation}
+Choosing $n^{-1/(2(s - 1) + d)}\leq r(n) \leq n^{-4/((2+d)(4s + d))}$, we may therefore apply our conclusions in Step 2; namely, that for every possible choice of $f \in W^{s,2}(\Xset;R)$ there exists a good set $\mathcal{E}_f \subseteq \Xset^n$ with $\Pbb(\mathcal{E}_f) \geq 1 - c/b$ such that each of \eqref{eqn:continuous_to_discrete_sobolev_norm}, \eqref{eqn:eigenvalue_tail_bound}, and \eqref{eqn:l2_to_empirical_norm_sobolev} hold for all $X \subseteq \mathcal{E}_f$.
+The choice $\kappa = (nR^2)^{2d/(4s + d)}$ balances the squared bias and variance terms on the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius}, and we have that for all $X \subseteq \mathcal{E}_f$
+\begin{align*}
+2b\sqrt{\frac{2\kappa}{n^2}} + \frac{f^T L^s f}{n\lambda_{\kappa}^s} & \leq 2bR^{2d/(4s + d)}n^{-4s/(4s+d)} + cbR^2\frac{n^{s} r^{s(d + 2)}}{\lambda_{\kappa}^s} \tag{by \eqref{eqn:continuous_to_discrete_sobolev_norm}} \\
+& \leq cbR^{2d/(4s + d)}n^{-4s/(4s+d)} \tag{by \eqref{eqn:eigenvalue_tail_bound}} \\
+& \leq \frac{1}{b}\norm{f}_{\Leb^2(\Xset)}^2 \\
+& \leq \frac{1}{n}\sum_{i = 1} \bigl[f(x_i)\bigr]^2. \tag{by \eqref{eqn:l2_to_empirical_norm_sobolev}}
+\end{align*}
+where the last two inequalities follow for a suitably large choice of $c_1$ in~\eqref{eqn:critical_radius_1}.
 We conclude that for all $X \subseteq \mathcal{E}_f$, the inequality \eqref{eqn:fixed_graph_testing_critical_radius} is satisfied with respect to $\beta = (f(x_1),\ldots,f(x_n))$ and $G = G_{n,r}$.
@@
 \begin{equation*}
-\sup_{\substack{f \in W^{1,2}(\mathcal{X;R}) \\ \norm{f}_{\Leb^2} \geq \epsilon}}\mathbb{E}_{f}(1 - \phi_{\spec}) \leq \sup_{\substack{f \in W^{1,2}(\mathcal{X;R}) \\ \norm{f}_{\Leb^2} \geq \epsilon}} \mathbb{E}\bigl[\mathbb{E}_{\beta}(1 - \phi_{\spec}|X \in \mathcal{E}_f)\bigr] + o(1) \leq \frac{3}{b} + o(1).
-\end{equation*}
+\sup_{\substack{f \in W^{s,2}(\Xset;R), \\ \norm{f}_{\Leb^2(\Xset)} \geq \epsilon}}\mathbb{E}_{f}(1 - \phi_{\spec}) \leq \sup_{\substack{f \in W^{s,2}(\Xset;R), \\ \norm{f}_{\Leb^2(\Xset)} \geq \epsilon}} \mathbb{E}\bigl[\mathbb{E}_{\beta}(1 - \phi_{\spec}|X \in \mathcal{E}_f)\bigr] + \frac{c}{b} \leq \frac{3 + c}{b},
+\end{equation*}
+completing the proof of Theorem~\ref{thm:sobolev_testing_rate} upon proper choice of constants $c_1,c_2$ in that Theorem.
+
+\clearpage
 
@@
 The proof of~\eqref{eqn:spectral_similarity} relies heavily on bounding the transport distance between $X$ and $\overline{X}$, i.e.
-\eqref{eqn:transport_distance}.
+\eqref{eqn:slepcev_transport_distance}.
 Then the following technical Lemma along with \eqref{eqn:spectral_similarity} will imply \eqref{eqn:pi_max}.
@@
 
-Here we give in full detail the proof of \eqref{eqn:spectral_similarity} and Lemma~\ref{lem:pi_max_pf_1}.
+Here we give in full detail the proofs of Lemma~\ref{lem:roughness_functional_expectation_sobolev}, \eqref{eqn:spectral_similarity}, and Lemma~\ref{lem:pi_max_pf_1}.
+
+\subsection{Proof of Lemma~\ref{lem:roughness_functional_expectation_sobolev}}
+To simplify exposition, we introduce the iterated difference operator, defined recursively as
+\begin{equation*}
+D_{jk}f(x) = (D_{k}f(x_j) - D_{k}f(x))\frac{K_r(x_j,x)}{r^d},~~ D_jf(x) = (f(x_j) - f(x))\frac{K_r(x_j,x)}{r^d}~~ \textrm{for $j \in [n], k \in [n]^q$}
+\end{equation*}
+We will also use the notation $d_jf(x) := (f(x_j) - f(x))$. We split our analysis into cases based on whether $s$ is even or odd.
+
+\subsubsection{Case 1: $s$ is even.}
+When $s$ is even, letting $q = s/2$ we have the decomposition
+\begin{equation}
+\label{eqn:continuous_to_discrete_sobolev_norm_pf1}
+f^T L^s f =  r^{ds} \cdot \sum_{i = 1}^{n} \sum_{k \in [n]^q} \sum_{\ell \in [n]^q} D_kf(x_i) D_{\ell}f(x_i). 
+\end{equation}
+
+For given index vectors $k,\ell \in [n]^q$ and indices $i,j$, let $I = \abs{k \cup \ell \cup i}$ be the total number of unique indices.
+We separate our analysis into cases based on the magnitude of $I$, specifically whether $I = s + 1$ (the leading terms where all indices are distinct) $I < s + 1$ (the terms where at least one index is repeated) and show that
+\begin{equation}
+\label{eqn:expected_difference_operators_1}
+\Ebb(D_kf(x_i) D_\ell f(x_i)) =
+\begin{cases*}
+O(r^{2s}) \cdot \norm{f}_{W^{s,2}(\Xset)}^2, & ~~\textrm{if $I = s + 1$} \\
+O(r^{2} r^{d(I - (s + 1))}) \cdot [f]_{W^{1,2}(\Xset)}^2, & ~~\textrm{if $I < s + 1$}~ 
+\end{cases*}
+\end{equation}
+We will prove~\eqref{eqn:expected_difference_operators_1} in Section~\ref{subsec:expected_difference_operators_pf}.
+First, we verify that~\eqref{eqn:continuous_to_discrete_sobolev_norm_pf1} and \eqref{eqn:expected_difference_operators_1} are together enough to show Lemma~\ref{lem:roughness_functional_expectation_sobolev} when $s$ is even.
+In the sum on the right hand side of~\eqref{eqn:continuous_to_discrete_sobolev_norm_pf1}, there are $O(n^{I})$ terms with exactly $I$ distinct indices.
+When $I < s + 1$, by~\eqref{eqn:expected_difference_operators_1} the total contribution of such terms to the sum is $O(n^{I}r^{d(I - 1) + 2}) \cdot [f]_{W^{1,2}(\Xset)}^2$.
+Since by assumption $r \geq n^{-1/d}$, this increases with $I$.
+Taking $I = s$ to be the largest integer less than $s + 1$, the contribution of these terms to the sum is therefore $O(n^sr^{d(s - 1) + 2}) \cdot [f]_{W^{1,2}(\Xset)}^2$ which in light of the restriction $r \geq n^{-1/(2(s - 1) + d)}$ is $O(n^{s+1}r^{s(d +2)}) \cdot [f]_{W^{1,2}(\Xset)}^2$.
+On the other hand when $I = s + 1$, by~\eqref{eqn:expected_difference_operators_1} we immediately have that the total contribution of these terms is $O(n^{s + 1}r^{2(s + d)}) \cdot \norm{f}_{W^{s,2}(\Xset)}$.
+Therefore,
+\begin{equation*}
+\Ebb(f^T L^s f) = O(n^{s+1}r^{s(d+2)}) \cdot \norm{f}_{W^{s,2}(\Xset)}.
+\end{equation*}
+
+\subsubsection{Case 2: $s$ is odd.}
+When $s$ is odd, letting $q = (s - 1)/2$ we have
+\begin{equation}
+\label{eqn:roughness_functional_representation_odd}
+f^T L^s f =  r^{ds} \cdot \sum_{i,j = 1}^{n} \sum_{k \in [n]^q} \sum_{\ell \in [n]^q} \bigl(d_jD_kf(x_i)\bigr) \cdot  \bigl(d_jD_{\ell}f(x_i)\bigr) \cdot K_r(x_i,x_j).
+\end{equation}
+For given index vectors $k,\ell \in [n]^q$ and indices $i,j \in [n]$, let $I = \abs{k \cup \ell \cup i \cup j}$ be the total number of unique indices.
+Similar to the case when $s$ is even, we show that
+\begin{equation}
+\label{eqn:expected_difference_operators_2}
+\Ebb(d_iD_kf(x_j) d_iD_\ell f(x_j)) =
+\begin{cases*}
+O(r^{2s}) \cdot \norm{f}_{W_d^{s,2}(\Xset)}^2, & ~~\textrm{if $I = s + 1$} \\
+O(r^{2} \cdot r^{d(I - (s + 1))}) \cdot [f]_{W_d^{1,2}(\Xset)}^2, & ~~\textrm{if $I < s + 1$}~ 
+\end{cases*}
+\end{equation}
+Then Lemma~\ref{lem:roughness_functional_expectation_sobolev} follows from similar reasoning to the case where $s$ was even.
+
+\subsubsection{Proof of~\eqref{eqn:expected_difference_operators_1} and~\eqref{eqn:expected_difference_operators_2}}
+\label{subsec:expected_difference_operators_pf}
+
+Note that if $f$ is constant almost everywhere, the claim is immediate as $D_kf(x_i) = 0$ with probability one. Otherwise $[f]_{W^{1,2}(\Xset)} > 0$, which we shall assume in what follows.
+
+Let $\delta = \min\{r^{2s},1\}\cdot[f]_{W^{1,2}(\Xset)} > 0$.
+Our analysis will make heavy use of Taylor expansions, and we therefore would like to show that there exists some smooth $g \in C^{\infty}(\Rd)$ such that
+\begin{equation*}
+\Bigl|\Ebb[D_kf(x_i) D_{\ell} f(x_i)] - \Ebb[D_kg(x_i) D_\ell g(x_i)]\Bigr| <  \delta,
+\end{equation*}
+and additionally $[g]_{W^{\ell,2}(\Rd)} \leq c [f]_{W^{\ell,2}(\Rd)}$ for each $\ell = 0,\ldots,s$.
+Then if \eqref{eqn:expected_difference_operators_1} and \eqref{eqn:expected_difference_operators_2} hold with respect to $g$, they hold (up to constants) with respect to $f$ as well.
+
+To construct such a $g$, we first take an extension of $f$ be defined on all of $\Rd$, and then mollify.
+Since $\Xset$ is a Lipschitz domain, there exists an extension \textcolor{red}{citation} $\wt{f}$ of $f$ compactly supported on $\Rd$ such that $\wt{f} = f$ a.e.
+on $\Xset$, and $[\wt{f}]_{W^{\ell,2}(\Rd)} \leq c [f]_{W^{\ell,2}(\Xset)}$ for each $\ell = 0,\ldots,s$.
+Since $\wt{f} = f$ a.e.
+on $\Xset$, the expected difference operators satisfy $\Ebb[D_k\wt{f}(x_i) D_\ell \wt{f}(x_i)] = \Ebb[D_kf(x_i) D_\ell f(x_i)]$.
+
+Now, since $\wt{f} \in W^{s,2}(\Rd)$, there exists a sequence $(g_m) \subset C^{\infty}(\Rd)$ such that $\norm{g_m - \wt{f}}_{W^{s,2}(\Rd)} \to 0$. On the one hand, by the Cauchy-Schwarz inequality
+\begin{equation*}
+\abs{\Ebb\bigl[D_k\wt{f}(x_i) D_{\ell} \wt{f}(x_i)\bigr] - \Ebb\bigl[D_kg_m(x_i) D_\ell g_m(x_i)\bigr]} \leq \frac{c}{r^{sd}} \cdot \norm{f - g_m}_{\Leb^2(\Rd)}
+\end{equation*}
+and taking $m$ to be sufficiently large, we can make the right hand side less than $\delta$.
+On the other hand, since $\norm{g_m - \wt{f}}_{W^{s,2}(\Rd)} \to 0$, there exists $m$ sufficiently large such that $[g_m]_{W^{s,2}(\Rd)}$ is at most say $2 [\wt{f}]_{W^{s,2}(\Rd)}$.
+We therefore take $g = g_m$ for $m$ large enough to satisfy both conditions.
+
+Our task is now to prove that \eqref{eqn:expected_difference_operators_1} and \eqref{eqn:expected_difference_operators_2} hold with respect to $g$.
+We first prove the desired bounds in the case when some indices are repeated, and then the desired bounds in the case when all indices are distinct.
+
+\paragraph{Repeated indices.}
+
+Since the proofs of~\eqref{eqn:expected_difference_operators_1} and~\eqref{eqn:expected_difference_operators_2} are essentially the same for the case where some index is repeated, we will assume without loss of generality that $s$ is even.
+Let $k,\ell \in [n]^q$ be index vectors for $q = s/2$.
+
+When at least one index is repeated, we obtain a sufficient upper bound by reducing the problem of upper bounding the iterated difference operator to that of upper bounding a single difference operator.
+Letting $k = (k_1,\ldots,k_q)$, we can show by induction that the absolute value of the iterated difference operator $\abs{D_kg(x_i)}$ is upper bounded by
+\begin{equation*}
+\abs{D_kg(x_i)} \leq \left(\frac{2K_{\max}}{r^d}\right)^{q-1} \sum_{h \in k \cup i} \abs{D_{k_q}g(x_h)} \cdot \1\{G_{n,r}[X_{k \cup i}]~\textrm{is a connected graph} \}.
+\end{equation*}
+Therefore,
+\begin{align}
+\abs{D_kg(x_i)} \cdot \abs{D_{\ell}g(x_i)} & \leq \left(\frac{2K_{\max}}{r^d}\right)^{2(q - 1)} \sum_{h,j \in k \cup \ell \cup i} \abs{D_{k_q}g(x_h)} \cdot \abs{D_{\ell_q}g(x_j)} \cdot \1\{G_{n,r}[X_{k \cup i}], G_{n,r}[X_{\ell \cup i}]~\textrm{are connected graphs.} \} \nonumber \\
+& =  \left(\frac{2K_{\max}}{r^d}\right)^{2(q - 1)} \sum_{h,j \in k \cup \ell \cup i} \abs{D_{k_q}g(x_h)} \cdot \abs{D_{\ell_q}g(x_j)} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is a connected graph.} \} \label{eqn:expected_difference_operators_sobolev_pf0}
+\end{align}
+
+We now break our analysis into three cases, based on the number of distinct indices in $k_q,\ell_q,h,j$. In each case we will obtain the same rate
+\begin{equation*}
+\Ebb\Bigl[\abs{D_{k_q}g(x_h)} \cdot \abs{D_{\ell_q}g(x_j)}\Bigr] = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot [g]_{W^{1,2}(\Rd)}^2,
+\end{equation*}
+and plugging this back in to~\eqref{eqn:expected_difference_operators_sobolev_pf0} we have that for any $k, \ell \in [n]^q$
+\begin{equation*}
+\Ebb\Bigl[\bigl|D_{k}g(x_i)\bigr| \cdot \bigl|D_{\ell}g(x_i)\bigr|\Bigr] = O(r^{(\abs{k \cup \ell \cup i} - (2q + 1))d + 2}) \cdot [g]_{W^{1,2}(\Rd)}^2.
+\end{equation*}
+
+\textit{Case 1: Two distinct indices.}
+Let $k_q = \ell_q = i$, and $h = j$. Using the law of iterated expectation, we obtain
+\begin{align*}
+\Ebb \left[\bigl(D_{i}g(x_j)\bigr)^2 \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\right] &= \Ebb \left[\bigl(D_{i}g(x_j)\bigr)^2 \cdot \Pbb\bigl[\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \} |x_i,x_j\bigr]\right] \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 2)d}) \cdot \Ebb\left[\bigl(D_{i}g(x_j)\bigr)^2\right] \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 3)d}) \cdot \Ebb\left[\bigl(d_{i}g(x_j)\bigr)^2K_r(x_i,x_j)\right] \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot [g]_{W^{1,2}(\Rd)}^2
+\end{align*}
+where the last equality follows from Lemma~\ref{lem:expected_first_order_seminorm}.
+
+\textit{Case 2: Three distinct indices.}
+Let $k_q = \ell_q = i$, for some $i \neq j \neq h$. Using the law of iterated expectation, we obtain
+\begin{align*}
+\Ebb \Bigl[ & \abs{D_{i}g(x_j)} \cdot \abs{D_ig(x_h)} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\Bigr] = \nonumber \\
+& \Ebb\Bigl[\abs{D_{i}g(x_j)} \cdot \abs{D_ig(x_h)} \cdot \Pbb\bigl[\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \} |x_i,x_j,x_h\bigr]\Bigr] \nonumber \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 3)d}) \cdot \Ebb\Bigl[\abs{D_{i}g(x_j)} \cdot \abs{D_ig(x_h)} \Bigr] \nonumber \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2})\cdot[g]_{W^{1,2}(\Rd)}^2
+\end{align*}
+where the last equality follows from Lemma~\ref{lem:expected_first_order_seminorm_2}.
+
+\textit{Case 3: Four distinct indices.}
+Using the law of iterated expectation, we find that
+\begin{align*}
+\Ebb\Bigl[ &\abs{D_{k_q}g(x_i)}\cdot{\abs{D_{\ell_q}g(x_j)}} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\Bigr] \\
+& = \Ebb\Bigl[\abs{D_{k_q}g(x_i)}\cdot{\abs{D_{\ell_q}g(x_j)}} \cdot\Pbb\bigl[G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected}|x_i,x_j,x_{k_q},x_{\ell_q}\bigr]\Bigr] \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 4)d}) \cdot \Ebb\Bigl[\abs{D_{k_q}g(x_i)}\cdot{\abs{D_{\ell_q}g(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr] \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot [g]_{W^{1,2}(\Rd)}^2
+\end{align*}
+where the last equality follows from Lemma~\ref{lem:expected_first_order_seminorm_3}.
+
+\paragraph{All indices distinct.}
+
+We first show the desired result when $s$ is even, and then when $s$ is odd.
+
+\textit{Case 1: $s$ is even.}
+
+By Lemma~\ref{lem:leading_term_sobolev} there exists some $f_s \in \Leb^2(\Rd)$ which satisfies $\norm{f_s}_{\Leb^2(\Rd)} \leq c \norm{g}_{W^{s,2}(\Rd)}$ such that
+\begin{equation*}
+\Ebb\Bigl[\bigl(D_kg(x)\bigr)^2\Bigr] = r^{2s} \cdot \norm{f_s}_{\Leb^2(\Rd)}.
+\end{equation*} Therefore, by the law of iterated expectation along with this Lemma, 
+\begin{align*}
+\Ebb\bigl[D_kg(x_i)D_kg(x_j)\bigr] = \Ebb\Bigl[\bigl(\Ebb[D_kg(x_i)|x_i]\bigr)^2\Bigr] = r^{2s} \cdot \Ebb\Bigl[\bigl(f_s(x_i)\bigr)^2\Bigr] = r^{2s} \cdot \norm{g}_{W^{s,2}(\Rd)},
+\end{align*}
+proving the claimed result.
+
+\textit{Case 2: $s$ is odd.}
+
+By the law of iterated expectation, we have
+\begin{align*}
+\Ebb[d_iD_kg_m(x_j)d_iD_{\ell}g_m(x_j)K_r(x_i,x_j)] & = \Ebb\biggl[\Bigl(d_i\bigl(\Ebb(D_kf)\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \\
+& = \Ebb\biggl[\Bigl(d_i\bigl(r^{s - 1}\cdot f_{s - 1} + r^sf_s(x_j)\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr].
+\end{align*}
+where the latter equality follows from Lemma~\ref{lem:leading_term_sobolev}. Here $f_{s -1}$ and $f_s$ satisfy the conclusions of that Lemma, namely that $f_{s - 1}$ and $f_s \in C^{\infty}(\Rd)$, and
+\begin{equation*}
+\norm{f_{s - 1}}_{W^{1,2}(\Rd)}, \norm{f_s}_{\Leb^2(\Rd)} \leq c \norm{g}_{W^{s,2}(\Rd)}.
+\end{equation*}
+By the linearity of the difference operator $d_i$, we have
+\begin{align*}
+\Ebb\biggl[\Bigl(d_i\bigl(r^{s - 1}\cdot f_{s - 1} + r^s f_s(x_j)\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] & = \Ebb\biggl[\Bigl(r^{s - 1} \cdot d_if_{s - 1}(x_j) + r^s d_if_s(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \\
+& \leq 2r^{2(s - 1)} \Ebb\Bigl[\bigl(d_if_{s-1}(x_j)\bigr)^2K_r(x_i,x_j)\Bigr] + 2r^{2s}\Ebb\Bigl[\bigl(d_if_s(x_j)\bigr)^2\Bigr] \\
+& \leq 2r^{2(s - 1)} \Ebb\Bigl[\bigl(d_if_{s-1}(x_j)\bigr)^2K_r(x_i,x_j)\Bigr] + 4r^{2s} \norm{g}_{W^{s,2}(\Rd)}^2 \\
+& \leq r^{2s} \cdot c\norm{g}_{W^{s,2}(\Rd)}^2
+\end{align*}
+where the last inequality follows from Lemma~\ref{lem:expected_first_order_seminorm}.
+This concludes the proof of~\eqref{eqn:expected_difference_operators_1} and~\eqref{eqn:expected_difference_operators_2}.
 
@@
 
-Assuming~\eqref{eqn:transport_distance} holds and $r$ is chosen according to~\eqref{eqn:pi_max}, we have already shown that the upper bound in~\eqref{eqn:spectral_similarity} holds for $\delta = 0$ (see~\eqref{eqn:eigenvalue_tail_bound_pf1}).
+Assuming~\eqref{eqn:slepcev_transport_distance} holds and $r$ is chosen according to~\eqref{eqn:pi_max}, we have already shown that the upper bound in~\eqref{eqn:spectral_similarity} holds for $\delta = 0$ (see the proof of ~\eqref{eqn:eigenvalue_tail_bound}).
 Now we need only to show the lower bound, which we will do following these steps:
@@
 In other words, we show that if edges in $G$ can be mapped to short, non-overlapping paths in $\wt{G}$, then the Laplacian matrix $L$ cannot be much larger $\wt{L}$, in the usual sense of partial ordering of matrices.
-	\item Now, we treat the case of $G = G_{n,r}$ and $\wt{G} = \overline{G}$.
-We assume that~\eqref{eqn:transport_distance} holds, and exhibit such a mapping between edges in $G_{n,r}$ and paths in $\overline{G}$.
+	\item Then, we treat the case of $G = G_{n,r}$ and $\wt{G} = \overline{G}$.
+We assume that~\eqref{eqn:slepcev_transport_distance} holds, and exhibit such a mapping between edges in $G_{n,r}$ and paths in $\overline{G}$.
 The resulting Poincare inequality will imply~\eqref{eqn:spectral_similarity}.
 \end{enumerate}
-In step (ii) we assume that~\eqref{eqn:transport_distance} holds.
-Since \eqref{eqn:transport_distance} holds with probability at least $1 - c/n$, this will imply that~\eqref{eqn:spectral_similarity} holds with at least the same probability.
+In step (ii) we assume that~\eqref{eqn:slepcev_transport_distance} holds.
+Since \eqref{eqn:slepcev_transport_distance} holds with probability at least $1 - c/n$, this will imply that~\eqref{eqn:spectral_similarity} holds with at least the same probability.
 
@@
 \subsubsection{Step 2: Mapping $E_{G_{n,r}}$ to $\mathcal{P}_{\overline{G}}$}
-We assume there exists a mapping $\pi:\overline{X} \to X$ which satisfies \eqref{eqn:transport_distance}.
+We assume there exists a mapping $\pi:\overline{X} \to X$ which satisfies \eqref{eqn:slepcev_transport_distance}.
 We will apply the Poincare inequality to $\pi^{-1}(G_{n,r})$, the isomorphism of $G_{n,r}$ induced by the mapping of vertices $\pi^{-1}: X \to \overline{X}$.
@@
 \section{Auxiliary Results}
+
+\subsection{Results of Others}
+
+Let $\overline{X}$ denote the $n$ evenly spaced grid points on $[0,1]^d$; formally, letting $t = n^{1/d}$
+\begin{equation*}
+\overline{X} = \biggl\{\frac{1}{t}(k_1,\ldots,k_d): k \in [t]^d\biggr\}
+\end{equation*}
+
+\begin{theorem}[Theorem 1 of \textcolor{red}{Garcia-Trillos and Slepcev}]
+	\label{thm:slepcev_transport_distance}
+	With probability at least $1 - c/n$ there exists a bijection $\pi: \overline{X} \to X$ such that
+	\begin{equation}
+	\label{eqn:slepcev_transport_distance}
+	\max_{k \in [t]^d} \abs{\overline{x}_k - \pi(\overline{x}_k)} \leq c \left(\frac{\log n}{n}\right)^{1/d}
+	\end{equation}
+\end{theorem}
 
@@
 
-For $u \in W^{1,2}(\Rd)$ and $x \in \Rd$, write $\nabla u(x) = (D^{e_1}(x),\ldots,D^{e_d}(x)$, for the gradient of $u$.
-
-\begin{theorem}[\textcolor{red}{Evans} Chapter 5.8.1, Theorem 2]
-	\label{thm:evans_poincare}
-	There exists a constant $C$, depending only on $d$, such that
-	\begin{equation*}
-	\norm{u - u^r}_{L^2(B(x,r))} \leq Cr\norm{\nabla u}_{L^2(B(x,r))}
-	\end{equation*}
-\end{theorem}
-
-\begin{theorem}[Taylor expansion.]
-	\label{thm:taylor_expansion}
-	For any function $u \in C^1$, and any $x,y \in \Rd$, 
-	\begin{equation*}
-	u(y) - u(x) = \int_{0}^{1} \nabla(u(x + t(y - x))) \cdot (y - x) \,dt
-	\end{equation*}
-\end{theorem}
-
-\begin{lemma}
-	\label{lem:poincare_mollify}
-	For any function $g \in W^{1,2}(\Rd)$ compactly supported in a bounded open set $V \subset \Rd$, we have
-	\begin{equation}
-	\label{eqn:poincare_mollify}
-	\int_{\Rd} (g(x) - g^r(x))^2 \,dx \leq r^2 \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
-	\end{equation}
-\end{lemma}
-\begin{proof}
-	This Lemma is essentially a reproduction of part of the proof of the Rellich-Kondrachov Compactness Theorem from \textcolor{red}{Evans}.
-Note that it is sufficient to prove in the case when $g$ is smooth.
-To see this, for the moment assume \eqref{eqn:poincare_mollify} holds for all $u \in C^{\infty}(V)$, and let $g \in W^{1,2}(V)$.
-By Theorem \ref{thm:local_approx_smooth_functions}, we may take a sequence $g_m \in C^{\infty}(V)$ such that
-	\begin{equation*}
-	\norm{g - g_m}_{L^2(V)} \to 0, \quad \textrm{and} \quad \norm{\nabla g - \nabla g_m}_{L^2(V)} \to 0,
-	\end{equation*}
-	Then we have
-	\begin{align*}
-	\int_{\Rd} (g(x) - g^r(x))^2 \,dx & \leq \int_{\Rd} (g(x) - g_m(x))^2 \,dx + \int_{\Rd}(g_m(x) - g_m^r(x))^2 \,dx + \int_{\Rd}(g_m^r(x) - g^r(x))^2 \,dx \\
-	& \leq \int_{\Rd} (g(x) - g_m(x))^2 \,dx + r^2 \int_{\Rd} \norm{\nabla g_m(x)}^2 \,dx + \int_{\Rd}(g_m^r(x) - g^r(x))^2 \,dx
-	\end{align*}
-	and taking the limit as $m$ goes to infinity, the right hand side converges to $\int_{\Rd} \norm{\nabla g(x)}^2 \,dx$.
-
-	It remains to show \eqref{eqn:poincare_mollify} in the case where $g$ is smooth. In this case, we have
-	\begin{align*}
-	g^{r}(x) -  g(x) & = \frac{1}{r^d} \int_{B(x,r)} \kappa \left(\frac{x - z}{r}\right)\bigl(g(z) - g(x)\bigr) \,dz \\
-	& = \int_{B(0,1)} \kappa(y) \bigl(g(x - ry) - g(x)\bigr) \,dy \\
-	& = \int_{B(0,1)} \kappa(y) \int_{0}^{1} \frac{d}{dt} \bigl(g(x - try) \bigr) \,dt \,dy \\
-	& = -r \int_{B(0,1)} \kappa(y) \int_{0}^{1} \bigl(\nabla g (x - try) \bigr) \cdot y \,dt \,dy.
-	\end{align*} 
-	Therefore, by Jensen's and Cauchy-Schwarz inequalities, we have
-	\begin{align*}
-	\int_{\Rd} (g(x) - g^r(x))^2 \,dx & \leq r^2 \int_{\Rd} \int_{B(0,1)} \kappa(y) \int_{0}^{1} \norm{\nabla g (x - try)}^2 \norm{y}^2 \,dt \,dy \,dx \\
-	& \leq r^2 \int_{B(0,1)} \kappa(y) \int_{0}^{1} \int_{\Rd} \norm{\nabla g (x - try)}^2 \,dx \,dt \,dy \\
-	& = r^2 \int_{\Rd} \norm{\nabla g(z)}^2 \,dz
-	\end{align*}
-\end{proof}
-
-The following theorem is Theorem 1 in Section 5.3 of \textcolor{red}{Evans}.
-\begin{theorem}[Local approximation by smooth functions.]
-	\label{thm:local_approx_smooth_functions}
-	Assume $U$ is bounded, and $u \in W^{k,p}(U)$ for some $1 \leq p < \infty$. Then there exists functions $u_m \in C^{\infty}(U) \cap W^{k,p}(U)$ such that
-	\begin{equation*}
-	\norm{u_m - u}_{W^{k,p}(U)} \overset{m}{\to} 0.
-	\end{equation*}
-\end{theorem}
-
-\begin{lemma}
-	\label{lem:gradient_mollify_commute}
-	For any $u \in W^{1,2}(\Rd)$, we have
-	\begin{equation*}
-	\int_{\Reals^d} \norm{\nabla u^r(x)}_2^2 \,dx \leq \int_{\Reals^d} \norm{\nabla u(x)}_2^2 \,dx
-	\end{equation*}
-\end{lemma}
-\begin{proof}
-	Observe that $\norm{\nabla u(x)}_2^2 = \sum_{j = 1}^{d} (D^{e_j}u(x))^2 $. Therefore,
-	\begin{align*}
-	\int_{\Rd} \norm{\nabla u^r(x)}_2^2 \,dP(x) & = \sum_{j = 1}^{d} \int_{\Rd} (D^{e_j}u^r(x))^2 \,dx \\
-	& = \sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u)^r(x))^2 \,dx
-	\end{align*}
-	where the second equality follows from equation (1) in Section 5.3 of \textcolor{red}{Evans}. Then, for any $v \in L^2(\Rd)$, we have that
-	\begin{align*}
-	\abs{v^r(x)} & = \int_{\Rd} \kappa_r^{1/2}(x - y) \kappa_r^{1/2}(x - y) v(y) \,dy \\
-	& \leq \left(\int_{\Rd} \kappa_r^(x - y) \,dy\right)^{1/2} \left(\int_{\Rd} \kappa_r(x - y) v^2(y) \right)^{1/2} \\
-	& = \left(\int_{\Rd} \kappa_r(x - y) v^2(y) \right)^{1/2}
-	\end{align*}
-	and therefore
-	\begin{align*}
-	\int_{\Rd} (v^r(x))^2 \,dx & \leq \int_{\Rd} \int_{\Rd } \kappa_r(x - y) v^2(y) \,dy \,dx \\
-	& = \int_{\Rd} \int_{\Rd } \kappa_r(x - y) v^2(y) \,dy \,dx \\
-	& = \int_{\Rd} v^2(y) \,dy
-	\end{align*}
-
-	Applying this to $D^{e_j}u \in L^2(\Rd)$, we have that
-	\begin{equation*}
-	\sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u)^r(x))^2 \,dx \leq \sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u(x))^2 \,dx = \int_{\Rd} \norm{\nabla u(x)}^2 \,dx.
-	\end{equation*}
-\end{proof}
-
-\begin{lemma}
-	\label{lem:paley_zygmund}
-	Let $0 < p < q$, and let $Z$ be a non-negative random variable such that $\mathbb{E}(Z^q) < \infty$. Then, for all $0 \leq \overline{s} \leq 1$,
-	\begin{equation}
-	\label{eqn:paley_zygmund}
-	\mathbb{P}(Z > \overline{s} \mathbb{E}(Z^p)) \geq \left[(1 - \overline{s}^p) \frac{\mathbb{E}(Z^p)}{(\mathbb{E}(Z^q))^{p/q}}\right]^{\frac{q}{q - p}}
-	\end{equation}
-\end{lemma}
-
 \begin{lemma}[Poincare inequality for path graphs.]
@@
 
+\subsection{Integrals}
+
+\begin{lemma}
+	\label{lem:expected_first_order_seminorm}
+	Suppose $g \in C^{\infty}(\Rd)$, that $\abs{p(x)} \leq p_{\max}$ for all $x \in \Rd$, and that $K$ is a $2$nd order kernel. Then
+	\begin{equation*}
+	\Ebb[(g(x_j) - g(x_i))^2K_r(x_i,x_j)] \leq c K_{\max} p_{\max}^2 r^2 [g]_{W^{1,2}(\Rd)}^2
+	\end{equation*}
+	for a constant $c$ which depends only on $\Xset$ and $d$.
+\end{lemma}
+\begin{proof}
+	By the fundamental theorem of calculus we have for any $y,x \in \Rd$,
+	\begin{equation*}
+	g(y) - g(x) = \int_{0}^{1} \frac{d}{dt}\bigl[g(x + t(y - x))\bigr] \,dt = \int_{0}^{1} \dotp{\nabla(g(x + t(y - x)))}{y - x} \,dt
+	\end{equation*}
+	By the upper bound on $p$, we obtain
+	\begin{align*}
+	\Ebb[(g(x_j) - g(x_i))^2K_r(x_i,x_j)] & \leq p_{\max}^2 \int_{\Rd} \int_{\Rd} (g(y) - g(x))^2 K_r(y,x) \,dy \,dx\\
+	& = p_{\max}^2 \int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \dotp{\nabla(g(x + t(y - x)))}{y - x} \,dt\right)^2 K_r(y,x) \,dy \,dx \\
+	& \overset{(i)}{\leq} p_{\max}^2 \int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \norm{\nabla(g(x + t(y - x)))}\norm{y - x} \,dt\right)^2 K_r(y,x) \,dy \,dx \\
+	& \overset{(ii)}{\leq} p_{\max}^2 r^2\int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \norm{\nabla(g(x + t(y - x)))}\,dt\right)^2 K_r(y,x) \,dy \,dx \\
+	& \overset{(iii)}{\leq} p_{\max}^2 r^2\int_{\Rd} \int_{\Rd} \int_{0}^{1} \norm{\nabla(g(x + t(y - x)))}^2 \,dt K_r(y,x) \,dy \,dx \\
+	& \overset{(iv)}{\leq} p_{\max}^2 K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(x,r)} \norm{\nabla(g(x + t(y - x)))}^2 \,dy \,dt \,dx \\
+	& \overset{(v)}{\leq}  p_{\max}^2 K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(0,r)} \norm{\nabla(g(x + z))}^2  \,dz \,dt \,dx
+	\end{align*}
+	where $(i)$ follows by Cauchy-Schwarz, $(ii)$ follows since either $\norm{y - x} \leq r$ or $K_r(y,x) = 0$, $(iii)$ follows by Jensen's, $(iv)$ follows by the assumption $K \leq K_{\max}$ supported on $B(0,1)$, and $(v)$ follows from the change of variables $z = x + t(y - x)$.
+Finally, again using Fubini's Theorem, we have
+	\begin{align*}
+	K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(0,r)} \norm{\nabla(g(x + z))}^2  \,dz \,dt \,dx & = r^{2 - d}\int_{B(0,r)} \int_{0}^{1} \int_{\Rd} \norm{\nabla(g(x + z))}^2  \,dz \,dt \,dx \\
+	& = K_{\max} r^2 [g]_{W_d^{1,2}(\Rd)}.
+	\end{align*}
+\end{proof}
+
+\begin{lemma}
+	\label{lem:expected_first_order_seminorm_2}
+	Suppose $g \in C^{\infty}(\Rd)$, that $\abs{p(x)} \leq p_{\max}$ for all $x \in \Rd$, and that $K$ is a $2$nd order kernel. Then
+	\begin{equation*}
+	\Ebb\Bigl[\abs{D_ig(x_h)}\cdot\abs{D_ig(x_j)} \Bigr] \leq c K_{\max} p_{\max}^2 r^2 [g]_{W^{1,2}(\Rd)}^2
+	\end{equation*}
+	for a constant $c$ which depends only on $\Xset$ and $d$.
+\end{lemma}
+\begin{proof}
+	We rewrite $\Ebb\bigl[\abs{D_{i}g(x_j)} \cdot \abs{D_ig(x_h)} \bigr]$ as follows,
+	\begin{align*}
+	\Ebb\Bigl[\abs{D_{i}g(x_j)} \cdot \abs{D_ig(x_h)} \Bigr] & = \int \int \int \abs{g(z) - g(x)} \cdot \abs{g(z) - g(y)} K_r(z,y) K_r(z,x) \,dP(x) \,dP(y) \,dP(x) \\
+	& = \int \left[\int \abs{g(z) - g(x)} K_r(z,x) \,dP(x)\right]^2 \,dP(z) \\
+	& \leq p_{\max}^3 \int_{\Xset} \left[\int_{\Xset} \abs{g(z) - g(x)} K_r(z,x) \,dx\right]^2 \,dz
+	\end{align*}
+	Then we obtain
+	\begin{align*}
+	\int_{\Xset} \abs{g(z) - g(x)} K_r(z,x) \,dx & \leq \int_{\Rd} \abs{g(z) - g(x)} K_r(z,x) \,dx \\
+	& = \int_{\Rd} \abs{\int_{0}^{1} \dotp{\nabla g(x + t(z - x))}{z - x} \,dt} K_r(z,x) \,dx \\
+	& \leq \int_{\Rd} \int_{0}^{1} \norm{\nabla g(x + t(z - x))}\cdot\norm{z - x} \,dt K_r(z,x) \,dx \\
+	& \leq r \int_{\Rd} \int_{0}^{1} \norm{\nabla g(x + t(z - x))} \,dt K_r(z,x) \,dx \\
+	& \leq r \frac{K_{\max}}{r^d} \int_{B(z,r)} \int_{0}^{1} \norm{\nabla g(x + t(z - x))} \,dt  \,dx \\
+	& \leq r K_{\max} \int_{B(0,1)} \int_{0}^{1} \norm{\nabla g(x - try)} \,dt  \,dy,
+	\end{align*}
+	and as a result, 
+	\begin{equation*}
+	p_{\max}^3 \int_{\Xset} \left[\int_{\Xset} \abs{g(z) - g(x)} K_r(z,x) \,dx\right]^2 \,dz \leq c\cdot p_{\max}^3 r^2 K_{\max}^3 [f]_{W_d^{1,2}(\Rd)}^2 = O(r^2)\cdot [f]_{W_d^{1,2}(\Xset)}^2.
+	\end{equation*}
+\end{proof}
+
+\begin{lemma}
+	\label{lem:expected_first_order_seminorm_3}
+	Suppose $g \in C^{\infty}(\Rd)$, that $\abs{p(x)} \leq p_{\max}$ for all $x \in \Rd$, and that $K$ is a $2$nd order kernel. Then
+	\begin{equation*}
+	\Ebb\Bigl[\abs{D_{k_q}g(x_i)}\cdot{\abs{D_{\ell_q}g(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr] \leq c K_{\max} p_{\max}^2 r^{2 + d} [g]_{W^{1,2}(\Rd)}^2
+	\end{equation*}
+	for a constant $c$ which depends only on $\Xset$ and $d$.
+\end{lemma}
+\begin{proof}
+	We rewrite the expectation as an integral,
+	\begin{align*}
+	\Ebb\Bigl[& \abs{D_{k_q}g(x_i)}\cdot{\abs{D_{\ell_q}g(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr] \\
+	& \leq p_{\max}^4 \int_{\Xset^4} \abs{g(x) - g(y)} \cdot \abs{g(u) - g(v)} \cdot  K_r(x,y) K_r(u,v) \1\{\norm{y - v} \leq (2q + 1)r\} \,dy \,dx \,du \,dv
+	\end{align*}
+	By substituting $z_1 = (y - v)/r$, $z_2 = (u - v)/r$, and $z_3 = (x - y)/r = (x - v)/r + z_1$, we can simplify the integral in the previous display,
+	\begin{align*}
+	\int_{\Xset^4} & \abs{g(x) - g(y)} \cdot \abs{g(u) - g(v)} \cdot K_r(x,y) K_r(u,v) \1\{\norm{y - v} \leq (2q + 1)r\} \,dy \,dx \,du \,dv \\
+	& \leq K_{\max}^2 r^d \int_{\Xset} \int_{[B(0,1)]^3} \abs{g\bigl((z_3 + z_1)r + v\bigr) - g(z_1r + v)} \cdot \bigl|g(z_2r + v) - g(v)\bigr| \,dz_1 \,dz_2 \,dz_3 \,dv \\
+	& \leq  K_{\max}^2 r^{d + 2} \int_{[B(0,1)]^3} \int_{[0,1]^2} \int_{\Xset} \norm{\nabla g(t z_3 r + z_1r + v)} \cdot \norm{\nabla g(t z_2 r + v)} \,dv \,dt_1 \,dt_2 \,dz_1 \,dz_2 \,dz_3 \\
+	& \leq c \nu_d^3 K_{\max}^2 r^{d + 2} [f]_{W_{d}^{1,2}(\Xset)}^2.
+	\end{align*}
+\end{proof}
+
+\begin{lemma}
+	\label{lem:remainder_term}
+	Suppose $f \in \Leb^2(\Rd)$, and $h:\Rd \times \Rd \times [0,1] \to \Reals$ is uniformly bounded.
+Then, the function $g(x) = \int_{0}^{1} \int_{B(0,1)} f(x + aty) h(y,x, t) \,dy \,dt$ also belongs to $\Leb^2(\Rd)$, with norm
+	\begin{equation*}
+	\norm{g}_{\Leb^2(\Rd)} \leq \nu_d \cdot \norm{f}_{\Leb^2(\Rd)} \cdot \norm{h}_{\infty}
+	\end{equation*}
+\end{lemma}
+\begin{proof}
+	We compute the squared norm of $g$,
+	\begin{align*}
+	\norm{g}_{\Leb^2(\Rd)}^2 & = \int_{\Rd} \left(\int_{0}^{1} \int_{B(0,1)} f(x + aty) h(y,x,t) \,dt \,dy \right)^2 \,dx \\
+	& \leq \norm{h}_{\infty}^2 \int_{\Rd} \left(\int_{0}^{1} \int_{B(0,1)} f(x + aty) \,dt \,dy \right)^2 \,dx \\
+	& \leq \nu_d^2 \norm{h}_{\infty}^2 \int_{\Rd} \int_{0}^{1} \frac{1}{\nu_d}\int_{B(0,1)} f^2(x + aty) \,dt \,dy \,dx \tag{Jensen's inequality} \\
+	& = \nu_d^2 \norm{h}_{\infty}^2 \int_{0}^{1} \int_{B(0,1)} \frac{1}{\nu_d}\int_{\Rd}f^2(x + aty) \,dt \,dy \,dx \tag{Fubini's theorem} \\
+	& = \nu_d^2 \norm{h}_{\infty}^2 \norm{f}_{\Leb^2(\Rd)}^2.
+	\end{align*}
+\end{proof}
+
+\begin{lemma}
+	\label{lem:leading_term_sobolev}
+	Let $k \in [n]^q$ for some $q \geq 1$, and let $K_r$ be a second order kernel.
+For $s \geq 0$ an integer, let $p$ be a density satisfying $p \in C^{s - 1}(\Rd;p_{\max})$ for some $p_{\max} > 0$ if $s \geq 1$, and otherwise $p \in C^{0}(\Rd;p_{\max})$.
+Then for any $f \in C^{\infty}(\Rd)$, there exist functions $f_{\ell}, \ell = 2q,\ldots,s$ satisfying $f_{\ell} \in C^{\infty}(\Rd)$ and
+	\begin{equation*}
+	\norm{f_{\ell}}_{W^{s - \ell,2}(\Rd)} \leq c \norm{f}_{W^{s,2}(\Rd)}
+	\end{equation*}
+	for some constant $c$ which depends only on $d$, $\Xset$ and $p_{\max}$ such that
+	\begin{equation}
+	\label{eqn:leading_term_sobolev}
+	\Ebb(D_kf(x)) =
+	\begin{cases*}
+	\sum_{\ell = 2q}^{s} f_{\ell}(x) r^{\ell} ,~~& \textrm{if $2q < s$} \\
+	r^s \cdot f_{s}(x),~~& \textrm{if $2q \geq s$}
+	\end{cases*}
+	\end{equation}
+\end{lemma}
+\begin{proof}	
+	We proceed by induction on $q$.
+Note that without loss of generality, we will assume $p \in C^{\infty}(\Rd)$, i.e.
+$p$ is smooth.
+Otherwise we can simply mollify $p$, with neglible change to the Holder norm $\norm{p}_{C^{s-1}(\Xset)}$ or the expectation $\Ebb[D_kf(x)]$.
+
+	\paragraph{Base case.}
+	We begin with the base case of $q = 1$. When $s = 0$, we need to show that $\Ebb[D_kf(x)] = f_s(x)$ for some $\norm{f_s}_{\Leb^2(\Rd)} \leq c\norm{f}_{W^{s,2}(\Rd)}$. But
+	\begin{align*}
+	\Ebb\Bigl[D_kf(x)\Bigr] & = \int f(z) K_r(x,z) p(z) \,dz  - f(x) \\
+	& = \int f(yr + x) K(\norm{y}) p(yr + z) \,dy - f(x).
+	\end{align*}
+	and since
+	\begin{equation*}
+	\biggl\|\int f(yr + x) K(\norm{y}) p(yr + x) \,dy\biggr\|_{\Leb^2(\Rd)} \leq p_{\max} K_{\max} \norm{f}_{\Leb^2(\Rd)}
+	\end{equation*}
+	the claim follows.
+
+	Now, when $s \geq 1$ since $f$ is smooth it admits a Taylor expansion of the following form for all $x,z \in \Rd$:
+	\begin{equation*}
+	f(z) = \sum_{\abs{\alpha} < s} \frac{f^{(\alpha)}(x)}{\alpha!} (x - z)^{\alpha} + \frac{1}{(s - 1)!}\sum_{\abs{\alpha} = s} (x - z)^{\alpha} \int_{0}^{1}(1 - t)^{s - 1} f^{(\alpha)}(x + t(z - x)) \,dt
+	\end{equation*}
+	where $f^{(\alpha)} \in C^{\infty}(\Rd)$ additionally satisfies
+	\begin{equation*}
+	\norm{f^{(\alpha)}}_{W^{s - \abs{\alpha},2}(\Rd)} \leq \norm{f}_{W^{s,2}(\Rd)}
+	\end{equation*}
+	Replacing $f$ by its Taylor expansion inside the expected first order difference operator $\Ebb(D_kf(x))$ we have
+	\begin{equation}
+	\label{eqn:leading_term_sobolev_pf1}
+	\Ebb(D_kf(x)) = \sum_{1 < \abs{\alpha} < s} \frac{f^{(\alpha)}(x)}{\alpha!} E_{\alpha,x} + \frac{1}{(s - 1)!}\sum_{\abs{\alpha} = s} \int_{0}^{1}(1 - t)^{s - 1} E_{\alpha,x,t}(f)  \,dt 
+	\end{equation}
+	where we use the notation $E_{\alpha,x} := \Ebb\left[(x - x_k)^{\alpha}K_r(x_k,x)\right]$ and  $E_{\alpha,x,t}(f) := \Ebb\bigl[f^{(\alpha)}(x + t(x_k - x)) (x_k - x)^{\alpha} K_r(x_k,x)\bigr]$.
+
+	 By a change of variables, we have
+	\begin{align*}
+	E_{\alpha,x,t}(f) & = \frac{1}{r^d}\int_{\Rd} f^{(\alpha)}\bigl(x + t(z - x)\bigr) (z - x)^{\alpha}  K\Bigl(\frac{\norm{z - x}}{r}\Bigr) p(z) \,dz \\
+	& = r^s \int_{\Rd} y^{\alpha} f^{(\alpha)}\bigl(x + tyr) K(\norm{y}) p(yr + x) \,dy,
+	\end{align*}
+	and as a result the remainder term in \eqref{eqn:leading_term_sobolev_pf1} reduces to
+	\begin{align*}
+	\frac{1}{(s - 1)!}\sum_{\abs{\alpha} = s} \int_{0}^{1}(1 - t)^{s - 1} E_{\alpha,x,t}(f)  \,dt & = \frac{r^s}{(s - 1)!}\sum_{\abs{\alpha} = s} \int_{0}^{1} \int_{\Rd} (1 - t)^{s - 1}  y^{\alpha} f^{(\alpha)}\bigl(x + tyr) K(\norm{y}) p(yr + x) \,dy \,dt \\
+	& =: r^s g_s(x) ~~\textrm{for $g_s \in C^{\infty}(\Rd)$} 
+	\end{align*}
+	where $\norm{g_s}_{\Leb^2(\Rd)} \leq c \norm{f}_{W^{s,2}(\Rd)}$ follows from Lemma~\ref{lem:remainder_term}.
+When $s = 1$, there are no $1 < \abs{\alpha} < s$, and so only the remainder term in \eqref{eqn:leading_term_sobolev_pf1} is non-zero; thus we have shown~\eqref{eqn:leading_term_sobolev} when $q = 1$ and $s = 1$.
+
+	When $s \geq 2$, we can analyze $E_{\alpha,x}$ using a Taylor expansion of $p$. For any $x,z \in \Rd$, we have
+	\begin{equation*}
+	p(z) = \sum_{\abs{\beta} < s  - 1} \frac{p^{(\beta)}(x)}{\beta!} (x - z)^{\beta} + \frac{1}{(s - 2)!}\sum_{\abs{\beta} = s - 1} (x - z)^{\beta} \int_{0}^{1}(1 - t)^{s - 2} p^{(\beta)}(x + t(z - x)) \,dt
+	\end{equation*}
+	and $p^{(\beta)} \in C^{\infty}(\Rd)$ additionally satisfies
+	\begin{equation*}
+	\norm{p^{(\beta)}}_{C^{s - \abs{\beta} - 1}(\Rd)} \leq p_{\max}.
+	\end{equation*}
+	Replacing $p$ by its Taylor expansion, the term $E_{\alpha,x}$ can be rewritten as
+	\begin{align*}
+	E_{\alpha,x} & = \int_{\Rd} (x - z)^{\alpha} K_r(x,z) p(z) \,dz \\
+	& =  \sum_{\abs{\beta} = 0}^{s - 2} \frac{p^{(\beta)}(x)}{\beta!} \int_{\Rd} (x - z)^{\alpha + \beta}K_r(x,z)\,dz ~~+ \\
+	& ~~~\frac{1}{(s - 2)!}\sum_{\abs{\beta} = s - 1} \int_{\Rd} \int_{0}^{1} (1 - t)^{s - 2} (x - z)^{\alpha + \beta - 1}  p^{(\beta)}(x + t(z - x)) K_r(x,z) \,dz \,dt \\
+	& = \sum_{\abs{\beta} = 0}^{s - 2} r^{\abs{\alpha} + \abs{\beta}} \frac{p^{(\beta)}(x)}{\beta!} \int_{\Rd} y^{\alpha + \beta}K(y)\,dy +~ \frac{r^{\abs{\alpha} + s - 1}}{(s - 2)!} \sum_{\abs{\beta} = s - 1} \int_{\Rd} \int_{0}^{1} (1 - t)^{s - 2} y^{\alpha + \beta - 1} p^{(\beta)}(x + ytr) K(\norm{y}) \,dy \,dt \\
+	& =: \sum_{\abs{\beta} = 0}^{s - 2} \1\{\abs{\alpha} + \abs{\beta} > 1\}r^{\abs{\alpha} + \abs{\beta}} p_{\abs{\beta}}(x) + r^s p_s(x) ~~\textrm{for $p_{\abs{\beta}} \in C^{\infty}(\Rd), \norm{p_{\abs{\beta}}}_{C^{s - \abs{\beta} - ,1}(\Rd)} \leq c \norm{p}_{C^{s - 1}(\Rd)}$},
+	\end{align*}
+	where the last line follows since $K$ is a 2nd order kernel. 
+	Substituting our expressions for $E_{\alpha,x}$ and $E_{\alpha,x,t}(f)$ back into~\eqref{eqn:leading_term_sobolev_pf1}, we obtain
+	\begin{align*}
+	\Ebb\bigl[D_kf(x)\big] & = \sum_{\abs{\alpha} = 1}^{s - 1} \sum_{\abs{\beta} = 0}^{s-2} \1\{\abs{\alpha} + \abs{\beta} > 1\} r^{\abs{\alpha} + \abs{\beta}} \frac{f^{(\alpha)}(x)p_{\abs{\beta}}(x)}{\alpha!} + r^s \biggl(\sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)p_s(x)}{\alpha!} + g_s(x)\biggr) \\
+	& = \sum_{\ell = 2}^{s - 1} r^{\ell} \sum_{ \substack{\abs{\alpha} + \abs{\beta} = \ell, \\ \abs{\beta} > 0} }   \frac{f^{(\alpha)}(x)p_{\abs{\beta}}(x)}{\alpha!} + r^s \biggl(\sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)p_s(x)}{\alpha!} + g_s(x)\biggr) \\
+	& =: \sum_{\ell = 2}^{s - 1} r^{\ell} f_{\ell}(x) + r^s f_s(x).
+	\end{align*}
+	It remains to show that $\norm{f_{\ell}}_{W^{s - \ell,2}(\Rd)} \leq c \norm{f}_{W^{s,2}(\Rd)}$ for each $\ell = 2,\ldots,s$.
+To see this, note that for $\abs{\alpha} + \abs{\beta} = \ell$ and $\min\{s - \abs{\alpha}, s - 1 - \abs{\beta}\} = m$, the inequality $s - \ell \leq m$ holds.
+Therefore,
+	\begin{equation*}
+	\norm{f^{(\alpha)}p_{\abs{\beta}}}_{W^{s - \ell,2}(\Rd)} \leq \norm{f^{(\alpha)}p_{\abs{\beta}}}_{W^{m,2}(\Rd)} \leq p_{\max} \norm{f^{(\alpha)}}_{W^{s - \abs{\alpha},2}(\Xset)} \leq p_{\max} \norm{f}_{W^{s,2}(\Rd)}
+	\end{equation*} 
+	which finishes the proof of Lemma~\ref{lem:leading_term_sobolev} when $q = 1$.
+
+	\paragraph{Induction Step.}
+
+	Now, we assume \eqref{eqn:leading_term_sobolev} holds for all $k \in [n]^q$, and prove the desired estimate on $\Ebb[D_{k}D_jf(x)]$ for each $j \in [n]$.
+We will build a proof piece-by-piece, depending on the relative size of $s$ and $q$.
+
+	If $s \leq 2$, by hypothesis there exists $f_s$ satisfying
+	\begin{equation*}
+	\norm{f_s}_{\Leb^2(\Rd)} \leq c \norm{f}_{W^{s,2}(\Rd)}
+	\end{equation*}
+	such that for any $z \in \Rd$,
+	\begin{equation*}
+	\Ebb\Bigl[D_jf(z)\Bigr] = r^s f_s(z).
+	\end{equation*}
+	Therefore by the law of iterated expectation,
+	\begin{equation*}
+	\Ebb\Bigl[D_kD_jf(x)\Bigr] = \Ebb\Bigl[D_k(\Ebb[D_jf])(x)\Bigr] = \Ebb\Bigl[D_k\bigl(r^s f_s\bigr)(x)\Bigr] = r^s \Ebb\Bigl[D_kf_s(x)\Bigr]
+	\end{equation*}
+	and again using the inductive hypothesis, we have that
+	\begin{equation}
+	\label{eqn:leading_term_sobolev_pf3}
+	\biggl\|\Ebb\Bigl[D_kf_s(x)\Bigr]\biggr\|_{\Leb^2(\Rd)} \leq c \norm{f_s}_{\Leb^2(\Rd)},
+	\end{equation}
+	establishing~\eqref{eqn:leading_term_sobolev} for all $q$ in the case when $s \leq 2$.
+
+	Otherwise, when $s \geq 3$ by hypothesis there exist functions $f_{\ell}$ for $\ell = 2,\ldots,s$ satisfying
+	\begin{equation*}
+	\norm{f_{\ell}}_{W^{s - \ell,2}(\Rd)} \leq c \norm{f}_{W^{s,2}(\Rd)}
+	\end{equation*}
+	such that for any $z \in \Rd$,
+	\begin{equation*}
+	\Ebb(D_jf(z)) = \sum_{\ell = 2}^{s} r^{\ell} f_{\ell}(z).
+	\end{equation*}
+	As a result, by the law of iterated expectation
+	\begin{align*}
+	\Ebb\Bigl[D_{k}D_jf(x)\Bigr] & = \Ebb\Bigl[D_k\bigl(\Ebb[D_jf]\bigr)(x)\Bigr] \\
+	& = \Ebb\biggl[D_k\Bigl(\sum_{\ell = 2}^{s - 1} r^{\ell} f_{\ell} + r^s f_s\Bigr)(x)\biggr] \\
+	& = \sum_{\ell = 2}^{s - 1} r^{\ell} \cdot \Ebb\Bigl[D_kf_{\ell}(x)\Bigr] + r^s \Ebb\Bigl[D_kf_s(x)\Bigr].
+	\end{align*}
+	Recalling that $\Ebb\Bigl[D_kf_s(x)\Bigr]$ satisfies~\eqref{eqn:leading_term_sobolev_pf3}, we now apply the inductive hypothesis to $\Ebb(D_kf_{\ell}(x))$ for each $\ell = 2,\ldots,s-1$, to prove~\eqref{eqn:leading_term_sobolev}.
+
+	First we consider the case when $2(q + 1) \geq s$.
+Note that $\ell \geq 2$ implies $2q \geq s - \ell$.
+Therefore by hypothesis, for each $\ell = 2,\ldots,s - 1$ the expectation $\Ebb[D_kf_{\ell}(x)] = r^{s - \ell}f_{\ell,s}(x)$ for some $f_{\ell,s} \in \Leb^2(\Rd)$ satisfying
+	\begin{equation}
+	\label{eqn:leading_term_sobolev_pf4}
+	\norm{f_{\ell,s}}_{\Leb^2(\Rd)} \leq c\norm{f_{\ell}}_{W^{s - \ell,2}(\Rd)} \leq c\norm{f}_{W^{s,2}(\Rd)}
+	\end{equation}
+	and as a result
+	\begin{equation*}
+	\sum_{\ell = 2}^{s - 1} r^{\ell} \cdot \Ebb(D_kf_{\ell}(x)) = r^s \sum_{\ell = 2}^{s - 1} f_{\ell,s}(x)
+	\end{equation*}
+	establishing that the second part of~\eqref{eqn:leading_term_sobolev} holds for all $q$ and all $s \leq 2(q + 1)$.
+
+	Otherwise $2(q + 1) < s - 1$.
+For each $\ell = 2,\ldots, s - 1$, if additionally  $2q \leq s - \ell - 1$, then by hypothesis $\Ebb(D_kf_{\ell}(x)) = \sum_{m = 2q}^{s - \ell - 1} r^{m} \cdot f_{\ell, \ell + m}(x) + r^{s - {\ell}} f_{\ell,s}$, and otherwise $\Ebb(D_kf_{\ell}(x)) = r^{s - \ell}f_{\ell,s}(x)$, where for each $\ell$ and $m$,
+	\begin{equation*}
+	\norm{f_{\ell,\ell + m}}_{W^{s - (\ell + m),2}(\Rd)} \leq c\norm{f_{\ell}}_{W^{s - \ell}} \leq c\norm{f}_{W^{s,2}(\Rd)}
+	\end{equation*}
+	and $f_{\ell,s}$ satisfies \eqref{eqn:leading_term_sobolev_pf4}. Therefore,
+	\begin{align*}
+	\sum_{\ell = 2}^{s - 1}  r^{\ell} \cdot \Ebb(D_kf_\ell(x)) & = \sum_{\ell = 2}^{s - 1 - 2q} r^{\ell} \cdot \left\{\sum_{m = 2q}^{s - \ell - 1} r^m \cdot f_{\ell,\ell + m}(x) + r^{s - \ell} \cdot f_{\ell,s}(x) \right\} + \sum_{\ell = s - 1 - 2q}^{s - 1} r^s \cdot f_{\ell,s}(x) \\
+	& = \sum_{\ell = 2}^{s - 1 - 2q} r^{\ell} \cdot \left\{\sum_{m = 2q}^{s - \ell - 1} r^m \cdot f_{\ell, \ell + m}(x)\right\}  + r^{s}\sum_{\ell = 2}^{s - 1}f_{\ell,s}(x) \\
+	& = \sum_{m = 2q}^{s - 3} \sum_{\ell = 2}^{s - m - 1} r^{m + \ell} \cdot f_{\ell, \ell + m}(x) + r^{s}\sum_{\ell = 2}^{s - 1}f_{\ell,s}(x).
+	\end{align*}
+	Rewriting the first sum in the final equation as a sum over $\ell + m = 2(q + 1),\ldots, s - 1$ establishes~\eqref{eqn:leading_term_sobolev}.
+\end{proof}
+
+The proof of Lemma~\ref{lem:empirical_norm_sobolev} relies on (a variant of) the Paley-Zygmund Inequality.
+\begin{lemma}
+	\label{lem:paley_zygmund}
+	Let $f$ satisfy the following moment inequality for some $b \geq 1$:
+	\begin{equation}
+	\label{eqn:paley_zygmund_1}
+	\Ebb\bigl[\norm{f}_n^4\bigr] \leq \left(1 + \frac{1}{b^2}\right)\cdot\Bigl(\Ebb\bigl[\norm{f}_n^2\bigr]\Bigr)^2.
+	\end{equation}
+	Then,
+	\begin{equation}
+	\label{eqn:paley_zygmund_2}
+	\mathbb{P}\left[\norm{f}_n^2 \geq \frac{1}{b} \Ebb\bigl[\norm{f}_n^2\bigr]\right] \geq 1 - \frac{5}{b}.
+	\end{equation}
+\end{lemma}
+\begin{proof}
+	Let $Z$ be a non-negative random variable such that $\mathbb{E}(Z^q) < \infty$. The Paley-Zygmund inequality says that for all $0 \leq \lambda \leq 1$,
+	\begin{equation}
+	\label{eqn:paley_zygmund_pf1}
+	\mathbb{P}(Z > \lambda \mathbb{E}(Z^p)) \geq \left[(1 - \lambda^p) \frac{\mathbb{E}(Z^p)}{(\mathbb{E}(Z^q))^{p/q}}\right]^{\frac{q}{q - p}}
+	\end{equation}
+	Applying~\eqref{eqn:paley_zygmund_pf1} with $Z = \norm{f}_n^2$, $p = 1$, $q = 2$ and $\lambda = \frac{1}{b}$, by assumption~\eqref{eqn:paley_zygmund_1} we have
+	\begin{equation*}
+	\mathbb{P}\Bigl(\norm{f}_n^2 > \frac{1}{b} \mathbb{E}[\norm{f}_n^2]\Bigr) \geq \Bigl(1 - \frac{1}{b}\Bigr)^2 \cdot  \frac{\bigl(\mathbb{E}[\norm{f}_n^2]\bigr)^2}{\mathbb{E}[\norm{f}_n^4]} \geq \frac{\Bigl(1 - \frac{2}{b}\Bigr)}{\Bigl(1 + \frac{1}{b^2}\Bigr)} \geq 1 - \frac{5}{b}.
+	\end{equation*}
+\end{proof}
+
 \end{document}

### 2020-01-25 notes/weekly_notes/1.16.20.NotesForWeek.tex
updated weekly notes and project level notes

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{\mathcal{L}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
	\title{Notes for Week of 1/9/20 - 1/16/20}
	\author{Alden Green}
	\date{\today}
	\maketitle

	Let $X = \{x_1,
	\ldots, x_n\}$ be a sample drawn i.i.d. from a distribution $P$ on $\Rd$,
	with density~$p$.  For a radius $r > 0$, we define $G_{n,r}=(V,E)$ to be the
	\emph{$r$-neighborhood graph} of $X$, an unweighted, undirected graph with
	vertices $V=X$, and an edge $(x_i,x_j) \in E$ if and only if $K_r(x_i,x_j) = \norm{x_i -x_j} \leq r$, where $\norm{\cdot}$ is the Euclidean norm. We denote by $A \in
	\Reals^{n \times n}$ the adjacency matrix, with entries $A_{uv} = 1$ if
	$(u,v) \in E$ and $0$ otherwise.  We also denote by $D$ the diagonal degree
	matrix, with entries $D_{uu} := \sum_{v \in V} A_{uv}$. The graph Laplacian is $L = D - A$, and we write its spectral decomposition as $L = V S V^T$.

	Suppose in addition to the random design points $X = \set{x_1,\ldots,x_n} \sim P$, we observe responses
	\begin{equation}
	\label{eqn:regression_known_variance}
	y_i = f(x_i) + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
	\end{equation} 
	To test whether $f = f_0$, we propose the following \emph{eigenvector projection} test statistic:
	\begin{equation}
	\label{eqn:graph_spectral_projections}
	T_{\mathrm{spec}} := \frac{1}{n} \sum_{k = 0}^{\kappa} \left(\sum_{i = 1}^{n} v_{k,i} y_i\right)^2
	\end{equation}
	where $v_k$ is the $k$th eigenvector of $L$ (ordered according to eigenvalues $s_1 \leq s_2 \leq \ldots \leq s_n$).

	The eigenvector projection test is minimax optimal over the balls in higher order Holder spaces $C_d^s(\Xset;L)$.

	\begin{theorem}
		\label{thm:higher_order_holder_testing_rate}
		Let $b \geq 1$ be a fixed constant, and let $d$ and $s$ be positive integers such that $d < 4s$. Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p \in C^{s-1}(\mathcal{X};R)$ bounded above and below by constants, i.e
		\begin{equation*}
		0 < p_{\min} < p(x) < p_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
		\end{equation*}
		Then the following statement holds: if the test $\phi_{\spec} = \1\{T_{\spec} \geq \tau \}$ is performed with parameter choices 
		\begin{equation*}
		n^{-1/(2(s - 1) + d)} \leq r(n) \leq n^{-4/((4s + d)(2+d))}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
		\end{equation*}
		then there exists constants $c_1,c_2$ which may depend on $d,R,$ and $s$ but are independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
		\begin{equation}
		\label{eqn:higher_order_holder_testing_rate}
		\epsilon^2 \geq c_1 \cdot b^2 \cdot n^{-4s/(4s + d)}
		\end{equation}
		the worst-case risk is upper bounded
		\begin{equation}
		\label{eqn:higher_order_holder_testing_rate_1}
		\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; C_d^{s}(\mathcal{X};R)) \leq \frac{c_2}{b}.
		\end{equation}
	\end{theorem}

\section{Proof}
Let $G = (V,E)$ be a graph over vertices $V = \set{v_1,\ldots,v_n}$, and let $\beta = (\beta_1,\ldots,\beta_n) \in \Reals^n$ be a signal over the vertices $V$. We observe responses $Y = (y_1,\ldots,y_n)$ according to the model
\begin{equation*}
y_i = \beta_i + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
\end{equation*}
Letting $L = VSV^T$ be the spectral decomposition of the Laplacian $L$ of $G$, our graph spectral test statistic is 
\begin{equation*}
T_{\spec} = \frac{1}{n}\sum_{k = 1}^{\kappa} \left(\sum_{i = 1}^{n} v_i y_i\right)^2
\end{equation*}
where $\kappa$ is a tuning parameter. The resulting test we will use is
\begin{equation*}
\phi_{\spec} = \1\{T_{\spec} \geq \frac{\kappa}{n} + t(b)\},~~\textrm{where}~ t(b) = b\sqrt{\frac{2\kappa}{n^2}}~~\textrm{for $b \geq 1$.}
\end{equation*}

Let $S_s(\beta;G)$ be a measure of smoothness the signal $\beta$ displays over the graph $G$, given by
\begin{equation*}
S_s(\beta;G) := \beta^T L^s \beta
\end{equation*}
In Lemma~\ref{lem:higher_order_fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}$. Our bound on the Type II error will be stated as a function of $S_2(\beta;G)$ as well as the $\kappa$th eigenvalue $s_{\kappa}$.

\begin{lemma}
	\label{lem:higher_order_fixed_graph_testing}
	Let $1 \leq \kappa \leq n$ be an integer.
	\begin{enumerate}
		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}$ is upper bounded
		\begin{equation}
		\label{eqn:higher_order_graph_spectral_type_I_error}
		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{1}{b^2}.
		\end{equation}
		\item \textbf{Type II error:} For any $b$ and $\beta$ such that
		\begin{equation}
		\label{eqn:higher_order_fixed_graph_testing_critical_radius}
		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_s(\beta;G)}{ns_{\kappa}^s}
		\end{equation}
		the Type II error of $\phi_{\spec}$ is upper bounded,
		\begin{equation}
		\label{eqn:higher_order_graph_spectral_type_II_error}
		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{3}{b}.
		\end{equation}
	\end{enumerate}
\end{lemma}

To prove Lemma~\ref{lem:higher_order_fixed_graph_testing} we will first compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}$, and then use Chebyshev's inequality to show \eqref{eqn:higher_order_graph_spectral_type_I_error} and \eqref{eqn:higher_order_graph_spectral_type_II_error}.

\vspace{.2 in}

\textit{Mean of $T_{\mathrm{spec}}$:} Using the notation $\dotp{v}{w} = \sum_{i = 1}^{n} v_iw_i$, we have
\begin{align*}
\Ebb(T_{\spec}) & = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{\beta}{v_k}^2 + \Ebb\bigl( \dotp{\varepsilon}{v_k}^2 + 2 \dotp{\varepsilon}{v_k} \dotp{\beta}{v_k}\bigr)\right) \\
& = \frac{\kappa}{n} + \frac{1}{n}\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2.
\end{align*}
When $\beta = 0$, this equals $\kappa/n$. Otherwise, we have the following lower bound:
\begin{align*}
\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2 & = \norm{\beta}_2^2 - \sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \\
& \geq \norm{\beta}_2^2 - \frac{1}{s_{\kappa}^s}\sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 s_k^s \\
& \geq \norm{\beta}_2^2 - \frac{S_s(\beta;G)}{s_{\kappa}^s},
\end{align*}
and therefore $\Ebb(T_{\spec}) \geq \kappa/n + n^{-1}(\norm{\beta}_2^2 - S_s(\beta;G)/s_{\kappa}^s)$.

\vspace{.2 in}

\textit{Variance of $T_{\mathrm{spec}}$:}
We write $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvectors $v_1,\ldots,v_{\kappa}$ as columns. Consequently,
\begin{align}
\Var(T_{\spec}) & = \frac{1}{n^2} \Var(y^T V_{\kappa} V_{\kappa}^T y) \\
& = \frac{1}{n^2} \Var((\beta + \varepsilon)^T V_{\kappa} V_{\kappa}^T (\beta + \varepsilon)) \\
& = \frac{1}{n^2} \Var(2 \beta^T V_{\kappa} V_{\kappa}^T \varepsilon + \varepsilon^T V_{\kappa} V_{\kappa}^T \varepsilon) \\
& \leq \frac{1}{n^2}(4 \beta^T V_{\kappa} V_{\kappa}^T \beta + 2\kappa)
\end{align}
where the last inequality follows from standard properties of the Gaussian distribution. We now move on to showing the desired inequalities \eqref{eqn:higher_order_graph_spectral_type_I_error} and \eqref{eqn:higher_order_graph_spectral_type_II_error}.

\vspace{.2 in}

\textit{Proof of~\eqref{eqn:higher_order_graph_spectral_type_I_error}:} By Chebyshev's inequality,
\begin{align*}
\Pbb_{\beta = 0}\bigl(T_{\spec} \geq \frac{\kappa}{n} + t(b)\bigr)
& \leq \Pbb_{\beta = 0}\bigl(\abs{T_{\spec} - \frac{\kappa}{n}} \geq t(b)\bigr) \\
& \leq \frac{\Var_{\beta = 0}(T_{\spec})}{t(b)^2} = \frac{1}{b^2}.
\end{align*}

\vspace{.2 in}

\textit{Proof of~\eqref{eqn:higher_order_graph_spectral_type_II_error}:} For simplicity, we introduce the notation
\begin{equation*}
\Delta = \frac{\norm{\beta}_2^2}{n} - \frac{S_s(\beta;G)}{ns_{\kappa}^s}.
\end{equation*}
Assumption~\eqref{eqn:higher_order_fixed_graph_testing_critical_radius} implies $\Delta \geq 2 t(b)$. Then another application of Chebyshev's inequality gives us
\begin{align*}
\Pbb_{\beta}\bigl(T_{\spec} \leq \frac{\kappa}{n} + t(b)\bigr) & = \Pbb_{\beta}\bigl(T_{\spec} - \Ebb_{\beta}(T_{\spec}) \leq t(b) - \Delta \bigr) \\
& \leq \Pbb_{\beta}\bigl(\abs{T_{\spec} - \Ebb_{\beta}(T_{\spec})} \leq \Delta - t(b) \bigr) \tag{since $\Delta \geq t(b)$}	\\
& \leq \frac{\Var_{\beta}(T_{\spec})}{(\Delta - t(b))^2} \\
& \leq 4\frac{\Var_{\beta}(T_{\spec})}{\Delta^2} \tag{since $\Delta \geq 2t(b)$} \\
& \leq 4\frac{2\kappa/n^2 + 4\beta^T V_{\kappa} V_{\kappa}^T \beta /n^2}{\Delta^2}.
\end{align*}
We handle each summand in the numerator separately. For the first term, since $\Delta \geq 2 t(b)$, we have
\begin{equation}
\label{eqn:higher_order_spectral_type_II_error_pf1}
\frac{2\kappa}{n^2\Delta^2} \leq \frac{1}{2b^2}.
\end{equation}

For the second term, noting that $\Delta = \beta^T V_{\kappa} V_{\kappa}^T \beta/n$, we have
\begin{align}
\frac{\beta^T V_{\kappa} V_{\kappa}^T \beta/n^2}{\Delta^2} & = \frac{1}{n\Delta} \nonumber \\
& \leq \frac{1}{2nt(b)} \nonumber \\
& = \frac{1}{2b\sqrt{2\kappa}}, \label{eqn:higher_order_spectral_type_II_error_pf2}
\end{align}
and combining~\eqref{eqn:higher_order_spectral_type_II_error_pf1} and~\eqref{eqn:higher_order_spectral_type_II_error_pf2} yields~\eqref{eqn:higher_order_graph_spectral_type_II_error}.

\subsubsection{Step 2: Bounding neighborhood graph functionals}

To make use of Lemma~\ref{lem:higher_order_fixed_graph_testing} we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:higher_order_fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$. In particular, we will show that for some constants $c_1,c_2,c_3,c_4$ which may depend on $L$, $d$ and $s$ but do not depend on $n$, $f$ or $b$, the following statements:
\begin{enumerate}[label=(E\arabic*)]
	\item 
	\label{event:higher_order_discrete_sobolev_norm}
	\textbf{Graph Sobolev norm:} For $f \in C_d^{s}(\mathcal{X};R)$, $p \in C_d^{s - 1}(\Xset;R)$, and $1 \geq r(n) \geq n^{-1/(2(s - 1) + d)}$,
	\begin{equation}
	\label{eqn:higher_order_continuous_to_discrete_sobolev_norm}
	S_s(f;G_{n,r}) \leq c_1 \cdot b \cdot n^{s + 1} r^{s(d + 2)} 
	\end{equation}
	\item 
	\label{event:higher_order_eigenvalue_tail_decay}
	\textbf{Eigenvalue tail bound:} For any $a >0$ and $(\log n/n)^{1/d} n^{a} \leq r \leq n^{-4/((2+d)(4s + d))}$,and for $\kappa = n^{2d/(4s + d)}$, 
	\begin{equation}
	\label{eqn:higher_order_eigenvalue_tail_bound}
	s_{\kappa} \geq c_2 \cdot n r^{d + 2} \kappa^{2/d}
	\end{equation}
	\item 
	\label{event:higher_order_l2_norm}
	\textbf{Empirical norm of $f$:} When $f \in C^{s}(\mathcal{X};R)$ and $\norm{f}_{\Leb^2} \geq c_3 \cdot b \cdot n^{-4s/(4s + d)}$,
	\begin{equation}
	\label{eqn:higher_order_l2_to_empirical_norm}
	\norm{f}_n^2 \geq \frac{1}{b} \cdot \norm{f}_{\Leb^2}^2
	\end{equation}
\end{enumerate}

each hold with probability at least $1 - c_4/b$ for sufficiently large $n$.

\paragraph{Proof of~\eqref{eqn:higher_order_continuous_to_discrete_sobolev_norm}:}

We will take $s$ to be even, as the proof when $s$ is odd follows essentially the same steps. To simplify exposition, we introduce the iterated difference operator, defined recursively as
\begin{equation*}
D_{jk}f(x) = (D_{k}f(x_j) - D_{k}f(x))\frac{K_r(x_j,x)}{r^d}~~\textrm{for $j \in [n], k \in [n]^q$}, \quad D_jf(x) = (f(x_j) - f(x))\frac{K_r(x_j,x)}{r^d}
\end{equation*}

Now when $s$ is even, letting $q = s/2$ we have the decomposition
\begin{equation}
\label{eqn:higher_order_continuous_to_discrete_sobolev_norm_pf1}
f^T L^s f =  \sum_{i = 1}^{n} \sum_{k \in [n]^q} \sum_{\ell \in [n]^q} r^{ds} D_kf(x_i) D_{\ell}f(x_i)
\end{equation}
For given index vectors $k,\ell$ and index $i$, let $I = \abs{k \cup \ell \cup i}$ be the total number of unique indices. We separate our analysis into cases based on the magnitude of $I$, specifically whether $I < s + 1$ or $I = s + 1$, and show that
\begin{equation}
\label{eqn:expected_difference_operators_1}
\Ebb(D_kf(x_i) D_\ell f(x_i)) =
\begin{cases*}
O(r^{2s}), & ~~\textrm{if $I = s + 1$} \\
O(r^{2} r^{d(I - (2q + 1))}), & ~~\textrm{otherwise}~ 
\end{cases*}
\end{equation}
uniformly over $f \in C^{s}(L)$. 
Before proving~\eqref{eqn:expected_difference_operators_1}, we verify that~\eqref{eqn:higher_order_continuous_to_discrete_sobolev_norm} is directly implied by~\eqref{eqn:expected_difference_operators_1}. In the sum on the right hand side of~\eqref{eqn:higher_order_continuous_to_discrete_sobolev_norm_pf1}, there are $O(n^{I})$ terms with exactly $I$ distinct indices. When $I < s + 1$, the total contribution of such terms to the sum is $O(n^{I}r^{d(I - 1) + 2})$. Since $r(n) \geq n^{-1/d}$, this increases with $I$. Taking $I = s$ to be the largest integer less than $s + 1$, the contribution of these terms to the sum is therefore $O(n^sr^{d(s - 1) + 2})$ which in light of the restriction $r \geq n^{-1/(2(s - 1) + d)}$ is $O(n^{s+1}r^{s(d +2)})$. On the other hand when $I = s + 1$, by~~\eqref{eqn:expected_difference_operators_1} we immediately have that the total contribution of these terms is $O(n^{s + 1}r^{2(s + d)})$. Therefore,
\begin{equation*}
\Ebb(f^T L^s f) = O(n^{s+1}r^{s(d+2)})
\end{equation*}
uniformly over $f \in C^s(L)$, and~\eqref{eqn:higher_order_continuous_to_discrete_sobolev_norm} by Markov's inequality.

Now we prove~\eqref{eqn:expected_difference_operators_1}. Since $f \in C_d^s(R) \subseteq C_d^1(R)$, using a first-order Taylor expansion of $f(x)$ we can show that for all index vectors $k,\ell \in [n]^q$ and indices $i \in [n]$, the product of iterated difference operators $D_kf(x_i) D_{\ell}f(x_i)$ satisfies
\begin{equation*}
\abs{D_kf(x_i) D_{\ell}f(x_i)} \leq 4^{q} R^2 r^{2 - 2dq}
\end{equation*}
Moreover $D_kf(x_i) D_{\ell}f(x_i)$ will equal zero if there exists $x_j, j \in k \cup \ell \cup i$ such that
\begin{equation*}
\norm{x_j - x_h} > r,~~\textrm{for all $h \neq j \in k \cup \ell \cup i$}
\end{equation*}
Therefore $D_kf(x_i) D_{\ell}f(x_i)$ is nonzero with probability $O(r^{d(\abs{k \cup \ell \cup i} - 1)})$, which along with the boundedness of $D_kf(x_i) D_{\ell}f(x_i)$ implies the second upper bound in~\eqref{eqn:expected_difference_operators_1}.

To show the first upper bound in~\eqref{eqn:expected_difference_operators_1}, we apply the law of iterated expectation, 
\begin{align}
\Ebb\left[D_kf(x_i)D_{\ell}f(x_i)\right] & = \Ebb\left[\Ebb\left(D_kf(x)|x_i = x\right) \Ebb\left(D_{\ell}f(x)|x_i = x\right)\right] \nonumber \\
& = \Ebb\left[\Ebb\left(D_kf(x)|x_i = x\right)^2\right]. \label{eqn:expected_difference_operators_pf1}
\end{align}
By Lemma~\ref{lem:leading_term}, we have that $\Ebb(D_kf(x)) = O(r^s)$ for all values of $x$, implying the desired result.

\paragraph{Proof of~\eqref{eqn:higher_order_eigenvalue_tail_bound}:}

We prove~\eqref{eqn:higher_order_eigenvalue_tail_bound} by comparing $G_{n,r}$ to the tensor product of a $d$-dimensional lattice and a complete graph. The latter is a highly structured graph with known eigenvalues, which as we will see are sufficiently lower bounded for our purposes.

Let $\wt{r} = r/(3(2\sqrt{d} + 1)), M = (1/\wt{r})^d, N = n\wt{r}^d$. Assume without loss of generality that $M$ and $N$ are integers. Additionally, for $t = n^{1/d}$ and $m = M^{1/d}$ let 
\begin{equation*}
\overline{X} = \set{\frac{1}{t}(k_1,\ldots,k_d): k \in [t]^d},~~ \overline{Z} = \set{\frac{1}{m}(j_1,\ldots,j_d): j \in [m]^d}.
\end{equation*}
For a given $\overline{z}_j \in \overline{Z}$, we write $Q(z_j) = m^{-1}[j_1 - 1,j_1] \times \cdots \times m^{-1}[j_d - 1,j_d]$ for the cube of side length $1/m$ with $z_j$ at one corner.

Consider the graph $H = (\overline{X}, E_H)$, where $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$ if
\begin{equation*}
\textrm{there exists}~\ol{z}_i, \ol{z}_j \in \ol{Z}~\textrm{such that}~\ol{x}_k \in Q(\ol{z}_i),~ \ol{x}_\ell \in Q(\ol{z}_j),~\textrm{and}~\norm{i - j}_1 \leq 1.
\end{equation*}
On the one hand $H \cong \ol{G}^M_d \otimes K_N$ where $\ol{G}^M_d$ is the $d$-dimensional lattice on $M$ nodes, and $K_N$ is the complete graph on $N$ nodes. On the other hand, we now show that with high probability $G_{n,r} \succeq H$. If $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$, then there exist $\ol{z}_i, \ol{z}_j$ such that
\begin{equation*}
\norm{\ol{x}_k - \ol{x}_{\ell}}_2 \leq m^{-1} + \norm{\ol{x}_k - \ol{z}_{i}}_2 + \norm{\ol{x}_{\ell} - \ol{z}_{j}}_2 \leq \wt{r}(1 + 2\sqrt{d}) = r/3.
\end{equation*}

Assuming~\eqref{eqn:transport_distance} holds, if $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$, then for sufficiently large $n$
\begin{equation*}
\norm{\pi(\ol{x}_k) - \pi(\ol{x}_\ell)}_2 \leq 2 c \left(\frac{\log n}{n}\right)^{1/d} + \frac{r}{3} \leq r,
\end{equation*}
implying that $(\pi(\ol{x}_k), \pi(\ol{x}_{\ell})) \in E$. Therefore, $G_{n,r} \succeq \ol{G}^M_d \otimes K_N$ whenever~\eqref{eqn:transport_distance} holds.

The eigenvalues of lattices and complete graphs are known to satisfy, respectively
\begin{equation*}
\lambda_k(\ol{G}^{M}_d) \geq \frac{k^{2/d}}{M^{2/d}}~\textrm{for $k = 0,\ldots,M - 1$},~~ \textrm{and}~\lambda_{j}(K_N) \geq N\1\{j > 0\}~\textrm{for $j = 0,\ldots,N-1$.}
\end{equation*}
and by standard facts regarding the eigenvalues of tensor product graphs, we have that the spectrum $\Lambda(H)$ satisfies
\begin{equation*}
\Lambda(H) = \set{N\lambda_k(\ol{G}^{M}_d) + M\lambda_j(K_N): \textrm{for $k = 0,\ldots,M - 1$ and $j = 0,\ldots,N-1$}}
\end{equation*}
For all $j = 1,\ldots,N-1$, we have that $M\lambda_j(K_N) = MN = n$. Therefore,
\begin{align*}
\lambda_{\kappa}(H) & \geq \{n \wedge N\lambda_{\kappa}(\ol{G}^{M}_d)\} \\
& \geq \{n \wedge n\wt{r}^d\frac{\kappa^{2/d}}{M^{2/d}}\} \\
& \geq \{n \wedge (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d}\} \\
& \geq (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d},
\end{align*}
where the last inequality can be verified by a quick calculation in light of $\kappa = n^{2d/(4s + d)}$ and $r \leq n^{-4/((2+d)(4s + d))}$. Since we've already shown that $\lambda_{\kappa}(G_{n,r}) \geq \lambda_{\kappa}(H)$ when~\eqref{eqn:transport_distance} is satisfied, which happens probability $1 - o(n^{-1})$, this completes the proof of~\eqref{eqn:higher_order_eigenvalue_tail_bound}.

\paragraph{Proof of~\eqref{eqn:higher_order_l2_to_empirical_norm}:}

Let $Z = \frac{1}{n}\sum_{i = 1}^{n} f^2(x_i)$. We upper bound $\Ebb[Z^2]$,
\begin{align*}
\Ebb[Z^2] & \leq \Ebb(f^2(x_1))^2 + \frac{1}{n}\Ebb(f^4(x_1)) \\
& \leq \Ebb(f^2(x_1))^2 + \frac{R^2}{n}\Ebb(f^2(x_1)) \\
& \overset{(i)}{\leq} \Ebb(f^2(x_1))^2\left(1 + \frac{R^2}{n/\Ebb(f^2(x_1))}\right) \\
& \overset{(ii)}{\leq} \Ebb(f^2(x_1))^2\left(1 + \frac{R^2}{c_3^2 b^2 n^{d/(4s + d)}}\right)
\end{align*}
where $(i)$ follows since $f \in C_d^s(\Xset;R)$ implies $\abs{f(x)} \leq R$, and $(ii)$ follows by assumption. The statement then follows by the Paley-Zygmund inequqality.

\subsubsection{Step 3: Conclusion}

We note that for all possible values of $X \in \Xset^n$, under the null hypothesis $f = f_0 = 0$ and therefore $\beta = (f(x_1),\ldots,f(x_n)) = 0$ as well. Therefore by~\eqref{eqn:higher_order_graph_spectral_type_I_error}, we have the following bound on Type I error:
\begin{equation}
\Ebb_{f_0}(\phi_{\mathrm{spec}}) = \mathbb{E}(\mathbb{E}_{\beta = 0}(\phi_{\spec}) | X) \leq \frac{1}{b^2}.
\end{equation}

Now, we bound Type II error under the assumption $f \in C_d^s(\mathcal{X};R), p \in C_d^{s - 1}(\Xset;R)$ uniformly bounded away from $0$ and $\infty$ over $\Xset$, and 
\begin{equation}
\label{eqn:higher_order_critical_radius_1}
\norm{f}_{\Leb^2}^2 \geq \epsilon^2 = c_3^2 \cdot b^2 \cdot n^{-4s/(4s + d)}.
\end{equation}
Choosing $n^{-1/(2(s - 1) + d)}\leq r(n) \leq n^{-4/((2+d)(4s + d))}$, we may therefore apply our conclusions in Step 2; namely, that for every possible choice of $f$ there exists a good set $\mathcal{E}_f \subseteq \Xset^n$ with $\Pbb(\mathcal{E}_f) \geq 1 - c_4/b$ such that each of \eqref{eqn:higher_order_continuous_to_discrete_sobolev_norm}, \eqref{eqn:higher_order_eigenvalue_tail_bound}, and \eqref{eqn:higher_order_l2_to_empirical_norm} hold for all $X \subseteq \mathcal{E}_f$. Choosing $\kappa = n^{2d/(4s + d)}$ to balance the squared bias and variance terms on the right hand side of~\eqref{eqn:higher_order_fixed_graph_testing_critical_radius}, we have that for all $X \subseteq \mathcal{E}_f$
\begin{align*}
2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_s(\beta;G_{n,r})}{ns_{\kappa}} & \leq 2bn^{-4s/(4s+d)} + c\cdot b \cdot \frac{n^{s} r^{s(d + 2)}}{s_{\kappa}^s} \tag{by \eqref{eqn:higher_order_continuous_to_discrete_sobolev_norm}} \\
& \leq 2bn^{-4s/(4s+d)} + c\cdot b \cdot\frac{1}{n^{4s/(4s+d)}} \tag{by \eqref{eqn:higher_order_eigenvalue_tail_bound}} \\
& \leq \frac{1}{b}\norm{f}_{\Leb^2} \\
& \leq \frac{1}{n}\sum_{i = 1} \beta_i^2. \tag{by \eqref{eqn:higher_order_l2_to_empirical_norm}}
\end{align*}
where the last two inequalities follow for a suitably large choice of $c_3$ in~\eqref{eqn:higher_order_critical_radius_1}.
We conclude that for all $X \subseteq \mathcal{E}_f$, the inequality \eqref{eqn:higher_order_fixed_graph_testing_critical_radius} is satisfied with respect to $\beta = (f(x_1),\ldots,f(x_n))$ and $G = G_{n,r}$. As a result the worst-case Type II error is bounded
\begin{equation*}
\sup_{\substack{f \in C_d^{s}(\mathcal{X;R}), \\ \norm{f}_{\Leb^2} \geq \epsilon}}\mathbb{E}_{f}(1 - \phi_{\spec}) \leq \sup_{\substack{f \in C_d^{s}(\mathcal{X;R}), \\ \norm{f}_{\Leb^2} \geq \epsilon}} \mathbb{E}\bigl[\mathbb{E}_{\beta}(1 - \phi_{\spec}|X \in \mathcal{E}_f)\bigr] + \frac{c_4}{b} \leq \frac{3 + c_4}{b},
\end{equation*}
completing the proof of Theorem~\ref{thm:higher_order_holder_testing_rate}.

\section{Supporting Results}

\begin{lemma}
	\label{lem:leading_term}
	Let $s$ be a positive integer, and $R \geq 0$. Suppose $f \in C^s(R)$, $p \in C^{s-1}(R)$, $k \in [n]^q$ for some $q \geq 1$, and that $K_r$ is a $2$nd-order kernel. Then if $2q \leq s - 1$,
	\begin{equation}
	\label{eqn:leading_term}
	\Ebb(D_kf(x)) = \sum_{\ell = 2q}^{s - 1} O(r^{\ell}) \cdot f_{\ell}(x) + O(r^s),
	\end{equation}
	for some $f_{\ell} \in C^{s - \ell}(L)$. If $2q \geq s$, $\Ebb(D_kf(x)) = O(r^s)$.  All $O(\cdot)$ terms may depend $L$ and $s$, but do not depend on $f$ or $x$.
\end{lemma}
\begin{proof}
	We will prove Lemma~\ref{lem:leading_term} in the case where $d = 1$. When $d \geq 2$, using multivariate Taylor expansions we find the same result, but with more notational overhead.

	We will prove by induction on $q$ in the case where $d = 1$. The When $q = 1$ and $k \in [n]$, by taking Taylor expansions of $f$ and $p$, we obtain
	\begin{align}
	\nonumber 
	\Ebb(D_kf(x)) & = \sum_{\ell = 1}^{s - 1} f^{(\ell)}(x) \int (z - x)^{\ell} K_r(z,x) p(z) dz + O(r^s) \\ 
	& = \sum_{\ell = 1}^{s - 1} \sum_{a = 0}^{s - 1} f^{(\ell)}(x) p^{(a)}(x) \underbrace{\int(z - x)^{\ell + a} K_r(z,x) dz}_{:=I_{t + a}} + O(r^s) \label{eqn:leading_term_pf1}
	\end{align}
	Since $K$ is a $2$nd-order kernel, $I_{1} = 0$ and $I_t = O(r^{t})$ for $t \geq 2$.  Additionally, when $\ell + a \leq s - 1$, we have that $f^{(\ell)}p^{(a)} \in C^{\min\{s - \ell, s - 1 - a\}} \subseteq C^{s - (\ell + a)}$, and $\abs{f^{(\ell)}p^{(a)}} \leq L^2$ for any $\ell$ and $a$. We can therefore simplify~\eqref{eqn:leading_term_pf1} by combining all terms where $\ell + a = t$, obtaining
	\begin{equation}
	\label{eqn:leading_term_pf2}
	\Ebb(D_kf(x)) = \sum_{t = 1}^{s - 1} f_t(x) I_t + O(r^s) = \sum_{t = 2}^{s - 1} f_t(x) I_t + O(r^s),
	\end{equation}
	which establishes~\eqref{eqn:leading_term} in the base case.

	Now, we assume \eqref{eqn:leading_term} holds for all $k \in [n]^q$, and prove the desired estimate on $\Ebb(D_{kj}f(x))$ for each $j \in [n]$. By the law of iterated expectation and~\eqref{eqn:leading_term_pf2},
	\begin{align*}
	\Ebb(D_{kj}f(x)) & = \Ebb(D_k(\Ebb(D_jf))(x)) \\
	& = \Ebb\left(D_k\left(\sum_{t = 2}^{s - 1} I_t f_t + O(r^s)\right)(x)\right) \\
	& = \sum_{t = 2}^{s - 1} I_t \cdot \Ebb(D_kf_t(x)) + O(r^s)
	\end{align*}
	where the second equality follows from the linearity and boundedness of $f \mapsto \Ebb(D_kf)$. We now apply the inductive hypothesis to $\Ebb(D_kf_t(x))$. If $2(q + 1) \geq s$, note that since $f_t \in C^{s - t}(L)$ for $t \geq 2$, we have by hypothesis $\Ebb(D_kf_t(x)) = O(r^{s - t})$. As a result
	\begin{equation*}
	\sum_{t = 2}^{s - 1} I_t \cdot \Ebb(D_kf_t(x)) = \sum_{t = 2}^{s - 1} I_t \cdot O(r^{s - t}) = O(r^s)
	\end{equation*} 
	Otherwise $2(q + 1) \leq s - 1$. For each $t = 2,\ldots, s - 1$, if additionally  $2q \leq s - t - 1$, then by hypothesis $\Ebb(D_kf_t(x)) = \sum_{\ell = 2q}^{s - t - 1} O(r^{\ell}) \cdot g_{\ell}(x) + O(r^{s - t})$ for some $g_{\ell} \in C^{s - t - \ell}(L)$, and otherwise $\Ebb(D_kf_t(x)) = O(r^{s - t})$. Therefore,
	\begin{align*}
	\sum_{t = 2}^{s - 1} I_t \cdot \Ebb(D_kf_t(x)) & = \sum_{t = 2}^{s - 1 - 2q} I_t \cdot \left\{\sum_{\ell = 2q}^{s - t - 1} O(r^{\ell}) \cdot g_{\ell}(x) + O(r^{s - t}) \right\} + \sum_{t = s - 1 - 2q}^{s - 1}I_{t} \cdot O(r^{s - t}) \\
	& = \sum_{t = 2}^{s - 1 - 2q} I_t \cdot \left\{\sum_{\ell = 2q}^{s - t - 1} O(r^{\ell}) \cdot g_{\ell}(x)\right\}  + O(r^s) \\
	& = \sum_{\ell = 2q}^{s - 3} \sum_{t = 2}^{s - \ell - 1} I_t \cdot O(r^{\ell}) \cdot g_{\ell}(x) + O(r^s).
	\end{align*}
	Noting that $g_{\ell} \in C^{s - (t + \ell)}(L)$ for some $\ell + t = 2(q + 1),\ldots,s - 1$, and $I_t \cdot O(r^{\ell}) = O(r^{t + \ell})$, we can rewrite the final equation as a sum over $\ell + t = 2(q + 1),\ldots,s - 1$, which proves~\eqref{eqn:leading_term}.
\end{proof}

\begin{theorem}[Theorem 1 of ``On the rate of Convergence of Empirical Measures in $\infty$-Transportation Distance'']
	Let $X = \set{x_1,\ldots,x_n} \sim P$, where $P$ is a distribution on $\Xset = [0,1]^d$ with density $p$ satisfying
	\begin{equation*}
	0 < p_{\min} < p(x) < p_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
	\end{equation*} 
	With probability at least $1 - o(n^{-1})$, there exists a bijection between grid points and data points $\pi: \overline{X} \to X$ such that
	\begin{equation}
	\label{eqn:transport_distance}
	\max_{k \in [t]^d} \abs{\overline{x}_k - \pi(\overline{x}_k)} \leq c \left(\frac{\log n}{n}\right)^{1/d}
	\end{equation}
\end{theorem}

\end{document}

### 2020-01-25 notes/weekly_notes/1.22.20.NotesForWeek.tex
updated weekly notes and project level notes

--- 2020-01-25 notes/graph_testing.tex
@@
 \usepackage{mathtools}
+\usepackage{physics}
 
@@
 \renewcommand{\bibsection}{\subsubsection*{\bibname}}
-
-\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
 
@@
 \newcommand{\myexp}[1]{\exp \{ #1 \}}
-\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
 \newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
@@
 \newcommand{\F}{\mathcal{F}}
-\newcommand{\D}{\mathcal{D}}
 \newcommand{\R}{\mathcal{R}}
@@
 \newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
-\newcommand{\gradient}{\mathcal{D}}
+% \newcommand{\gradient}{\nabla}
 \newcommand{\Hessian}{\mathcal{D}^2}
 \newcommand{\dotp}[2]{\langle #1, #2 \rangle}
+\newcommand{\deriv}{\mathcal{D}}
 
@@
 \newcommand{\spec}{\mathrm{spec}}
+\newcommand{\Unq}{\mathrm{Unq}}
 
@@
 \begin{document}
-\title{Graph Testing}
+\title{Notes for Week 1/15/20 - 1/22/20}
 \author{Alden Green}
@@
 
-This document details the current status of the graph testing project.
-We divide by section based on the testing problem under consideration.
-At the end of each section, we list some areas we are interested in investigating.
-All proofs are left until the end.
-
-We establish some notation that we will use throughout. Let $X = \{x_1,
+Let $X = \{x_1,
 \ldots, x_n\}$ be a sample drawn i.i.d. from a distribution $P$ on $\Rd$,
@@
 \emph{$r$-neighborhood graph} of $X$, an unweighted, undirected graph with
-vertices $V=X$, and an edge $(x_i,x_j) \in E$ if and only if $K_r(x_i,x_j) = \norm{x_i -x_j} \leq r$, where $\norm{\cdot}$ is the Euclidean norm. We denote by $A \in
+vertices $V=X$, and an edge $(x_i,x_j) \in E$ if and only if $K(\norm{x_i - x_j}) := \1\{\norm{x_i -x_j} \leq r\}$, where $\norm{\cdot}$ is the Euclidean norm. We denote by $A \in
 \Reals^{n \times n}$ the adjacency matrix, with entries $A_{uv} = 1$ if
@@
 
-\section{Regression goodness-of-fit testing with random design.}
-
-Let $P$ be a distribution with density $p$ supported on $\mathcal{X} \subseteq \Reals^d$. Suppose we observe random design points $X = \set{x_1,\ldots,x_n} \sim P$, and additionally responses
+Suppose in addition to the random design points $X = \set{x_1,\ldots,x_n} \sim P$, we observe responses
 \begin{equation}
@@
 \end{equation} 
-
-We wish to distinguish
-\begin{equation*}
-\mathbf{H}_0: f = f_0 := 0 \quad \textrm{vs} \quad \mathbf{H}_{\textrm{a}}: f \neq f_0
-\end{equation*}
-We will evaluate our performance using worst-case risk: for a given function class $\mathcal{H}$ and test function $\phi: \Reals^n \to \set{0,1}$, let
-\begin{equation*}
-\mathcal{R}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, f \neq f_0} \Ebb_f(1 - \phi).
-\end{equation*}
-The worst-case risk may be quite close to $1$ unless we enforce some separation between null and alternative spaces. A more realistic measure of performance is therefore
-\begin{equation*}
-\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, \norm{f - f_0}_{\Leb^2} \geq \epsilon} \Ebb_f(1 - \phi).
-\end{equation*}
-
-\subsection{Test Statistics.}
-We list the test statistics we use for the regression testing problem.
-
-\paragraph{Eigenvector projection test statistic.}
-Let $VSV^T$ be the spectral decomposition of the Laplacian matrix $L$ of the neighborhood graph $G_{n,r}$.
-To test whether $f = f_0$, we propose the following \emph{eigenvector projection} test statistic:
+To test whether $f = 0$, we propose the following \emph{eigenvector projection} test statistic:
 \begin{equation}
@@
 \end{equation}
-
-\subsection{Current Results.}
-
-In Theorem~\ref{thm:sobolev_testing_rate} we show that under some typical regularity conditions on $P$, the graph Laplacian eigenvector projection test $\phi_{\textrm{spec}} := \1\{T_{\mathrm{spec}} \geq \tau\}$ is a minimax optimal test over the Sobolev balls $W^{s,2}(\mathcal{X};R)$ whenever $4s > d$ and $\Xset = [0,1]^d$.
+where $v_k$ is the $k$th eigenvector of $L$ (ordered according to eigenvalues $s_1 \leq s_2 \leq \ldots \leq s_n$).
+
+The eigenvector projection test is minimax optimal over the balls in higher order Sobolev spaces $W_d^{s,2}(\Xset)$.
 
 \begin{theorem}
-	\label{thm:sobolev_testing_rate}
-	Let $b \geq 1$ and $R > 0$ be fixed constants.
-Suppose that $\Pbb$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p \in C^{s-1}(\Xset;p_{\max})$ for some $p_{\max} < \infty$, and further that $p(x)$ is bounded away from zero, i.e.
-there exists $p_{\min} > 0$ such that
-	\begin{equation*}
-	p_{\min} < p(x),~~ \textrm{for all $x \in \mathcal{X}$.}
-	\end{equation*}
-	Then the following statement holds: if the test $\phi_{\spec}$ is performed with parameter choices 
-	\begin{equation*}
-	\textrm{$K$ the uniform kernel,}~ n^{-1/(2(s-1) + d)} \leq r(n) \leq n^{-4/((4s + d)(2+d))}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
-	\end{equation*}
-	then there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
-	\begin{equation}
-	\label{eqn:sobolev_testing_rate}
-	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4s + d)}\} \cdot n^{-{4s}/(4s + d)}
+	\label{thm:higher_order_sobolev_testing_rate}
+	Let $b \geq 1$ be a fixed constant, and let $d$ and $s$ be positive integers such that $d < 4s$.
+Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p \in C^{s-1}(\mathcal{X};R)$ bounded above and below by constants, i.e
+	\begin{equation*}
+	0 < p_{\min} < p(x) < p_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
+	\end{equation*}
+	Then the following statement holds: if the test $\phi_{\spec} = \1\{T_{\spec} \geq \tau \}$ is performed with parameter choices 
+	\begin{equation*}
+	n^{-1/(2(s - 1) + d)} \leq r(n) \leq n^{-4/((4s + d)(2+d))}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	\end{equation*}
+	then there exists constants $c_1,c_2$ which may depend on $d$ and $s$ but are independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
+	\begin{equation}
+	\label{eqn:higher_order_sobolev_testing_rate}
+	\epsilon^2 \geq c_1 \cdot b^2 \cdot R^2 \cdot n^{-4s/(4s + d)}
 	\end{equation}
@@
 	\begin{equation}
-	\label{eqn:sobolev_testing_rate_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W^{2,s}(\mathcal{X};R)) \leq \frac{c_2}{b}.
+	\label{eqn:higher_order_sobolev_testing_rate_1}
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W_d^{s,2}(\mathcal{X};R)) \leq \frac{c_2}{b}.
 	\end{equation}
 \end{theorem}
-
-The restriction that $p$ be supported on the unit cube is mostly for convenience.
-If instead $p$ is supported on any compact set with Lipschitz boundary, the Theorem statement should hold (up to constants), with only a few modifications to our proofs.
-However, we do not work through the details.
-
-The restriction $4s > d$ is a fundamental consequence of the tightness of the Sobolev embedding theorem. When $4s \geq d$ the compact embedding
-\begin{equation*}
-W_d^{s,2}(\mathcal{X}) \subseteq \mathcal{L}_d^4(\mathcal{X}) 
-\end{equation*}
-does not hold (as it does when $d < 4s$). However, if we directly assume $f \in \mathcal{L}_d^4(\mathcal{X};R)$ (regardless of $d$) we obtain the following result.
-\begin{proposition}
-	\label{prop:L4_testing_rate}
-	If $f \in \Leb^4(\mathcal{X};R)$, there exists a constant $c$ such that if
-	\begin{equation}
-	\label{eqn:L4_testing_rate}
-	\epsilon^2 > b \cdot n^{-1/2}
-	\end{equation}
-	then the test
-	\begin{equation*}
-	\phi_{\mathrm{mean}} = \1\{\frac{1}{n}\sum_{i = 1}^{n} y_i^2 \geq 1\}
-	\end{equation*}
-	has worst-case risk
-	\begin{equation*}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}. 
-	\end{equation*}
-\end{proposition}
-Note that when $4s > d$ the critical radius~\eqref{eqn:sobolev_testing_rate} is smaller than \eqref{eqn:L4_testing_rate}.
-
-\subsection{Areas to Investigate}
-
-\begin{enumerate}[(i)]
-	\item When $d \geq 4$, is it true that there does not exist a uniformly consistent
-	test over the Sobolev ball $W_d^{1,2}(\mathcal{X};R)$?
-	\item Assume the distribution $P$ is supported on a manifold $\mathcal{M}$ of intrinsic dimension $s < d$.
-Does $\mathrm{\phi_{\mathrm{spec}}}$ display adaptivity to the intrinsic dimension of $\mathcal{M}$?
-	\item Assume that $f$ belongs to the Holder space $C_d^s(\mathcal{X})$.
-Moreover, suppose that instead of observing ${y_i}$ according to the regression testing model \eqref{eqn:regression_known_variance}, we observe
-	\begin{equation*}
-	y_i = f(x_i) + \sigma \varepsilon_i, ~\varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
-	\end{equation*}
-	where $\sigma > 0$ is unknown.
-When $d \geq 4$, what are the minimax regression testing rates over $C_d^1(\mathcal{X};R)$?
-Is the test $\phi_{\mathrm{spec}}$ minimax optimal, when the tuning parameters $r$ and $\kappa$ are appropriately chosen?
-\end{enumerate}
-
-\section{Two-sample density testing.}
-In the two-sample density testing problem, we observe independent samples $Z = z_1,\ldots,z_N \sim P$ and $Y = y_1,\ldots,y_M \sim Q$, where $P$ and $Q$ are distributions over $\Reals^d$ with densities $p$ and $q$, respectively, and $N \sim \textrm{Bin}(n,1/2)$.
-Our goal is to distinguish the hypotheses
-\begin{equation*}
-\mathbf{H}_0: P = Q \quad \textrm{vs} \quad \mathbf{H}_{\textrm{a}}: P \neq Q
-\end{equation*}
-and we again evaluate our performance using worst-case risk; letting $\phi:\Reals^{N + M} \to \{0,1\}$, 
-\begin{equation*}
-\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \inf_{p \in \mathcal{H}}\Ebb_{p,p}(\phi) + \sup_{\substack{p,q \in \mathcal{H} \\ \norm{p - q}_{\Leb^2} \geq \epsilon}} \Ebb_{p,q}(1 - \phi).
-\end{equation*}
-
-\subsection{Test statistics.}
-
-We suggest several two-sample test statistics.
-
-\paragraph{Eigenvector projection test statistic.}
-
-It is straightforward to adapt the test statistic $T_{\mathrm{spec}}$ to the two-sample testing problem.
-Concatenate the samples $Z$ and $Y$ in $X = (z_1,\ldots,z_N,y_1,\ldots,y_M)$, and define $T_{\mathrm{spec}}^{(2)}$ to be
-\begin{equation}
-\label{eqn:graph_spectral_projections_2}
-T_{\mathrm{spec}}^{(2)} := \frac{1}{n} \sum_{k = 0}^{\kappa} \left(\sum_{i = 1}^{n} v_i a_i\right)^2, ~~\textrm{where}~~ a = (\underbrace{N^{-1},\ldots,{N^{-1}}}_{\textrm{length } N},\underbrace{-M^{-1},\ldots,-M^{-1}}_{\textrm{length } M})
-\end{equation}
-
-For convenience, we state our following two test statistics with respect to the empirical norm $\norm{\theta}_n = n^{-1/2}\norm{\theta}_2$ for $\theta \in \Reals^n$.
-They will each depend on a tuning parameter $\lambda > 0$.
-\paragraph{Graph Sobolev IPM.}
-Letting $C_n := nr^{(d + 2)/2}$ and
-\begin{equation*}
-\Theta_{1,2} := \{\theta \in \Reals^n:~ C_n^{-1} \norm{B\theta}_2 \leq 1\} 
-\end{equation*}
-we define the \emph{graph Sobolev IPM} to be
-\begin{equation}
-\label{eqn:sobolev_IPM}
-T_{\textrm{sob}} := \sup_{\substack{\theta \in \Theta_{1,2} \\ \lambda \norm{\theta}_n \leq 1}} \abs{\frac{1}{n}\sum_{i = 1}^{n} a_i \theta_i}. 
-\end{equation}
-
-\paragraph{Graph Total Variation IPM.}
-Letting $C_n' := n^{2}r^{(d + 1)}$ and 
-\begin{equation*}
-\Theta_{1,1} := \{\theta \in \Reals^n:~ (C_n')^{-1} \norm{B\theta}_1 \leq 1\}
-\end{equation*}
-we define the \emph{graph Total Variation} IPM to be
-\begin{equation}
-\label{eqn:total_variation_IPM}
-T_{\mathrm{TV}} := \sup_{\substack{\theta \in \Theta_{1,1}, \\ \lambda \norm{\theta}_n \leq 1} } \abs{\frac{1}{n}\sum_{i = 1}^{n} a_i \theta_i}, \quad
-\end{equation}
-
-\subsection{Current results.}
-
-In Theorem~\ref{thm:twosample_sobolev_testing_rate} we show that under some typical regularity conditions on $P$, the test $\phi_{\textrm{spec}}^{(2)} := \1\{T_{\mathrm{spec}}^{(2)} \geq \tau\}$ is, up to log factors, a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};R)$ when $d = 1$.
-
-\begin{theorem}
-	\label{thm:twosample_sobolev_testing_rate}
-	Let $b \geq 1$ and $a > 0$ be fixed constants, and let $d = 1$.
-Suppose that $\mu = (P + Q)/2$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]$ with density functions $\rho$ bounded above and below by constants, i.e
-	\begin{equation*}
-	0 < \rho_{\min} < \rho(x) < \rho_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
-	\end{equation*}
-	Then the following statement holds: if the test $\phi_{\spec}^{(2)}$ is performed with parameter choices 
-	\begin{equation*}
-	r = \log^a n \cdot \left(\frac{\log n}{n}\right), ~\kappa = n^{2/5}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
-	\end{equation*}
-	then there exists a constant $c$ which may depend on $R,p_{\max},q_{\max},b$ and $a$ but is independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
-	\begin{equation}
-	\label{eqn:twosample_sobolev_testing_rate}
-	\epsilon^2 \geq c \cdot b \cdot n^{-4/5} (\log n)^{h(a,1)}
-	\end{equation}
-	the worst-case risk is upper bounded
-	\begin{equation}
-	\label{eqn:twosample_sobolev_testing_rate_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; \mathcal{W}^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}.
-	\end{equation}
-\end{theorem}
-
-We prove Theorem~\ref{thm:twosample_sobolev_testing_rate} by relating the density testing problem to a regression testing problem with a certain type of structured noise, and then proceeding along similar lines to the proof of Theorem~\ref{thm:sobolev_testing_rate}.
-To pursue this strategy, we require the eigenvectors to satisfy a certain type of incoherence condition; this is in constrast to the regression testing problem with known variance, where we did not require the eigenvectors to be smooth in any sense.
-
-\subsection{Areas to Investigate}
-
-\begin{enumerate}[(i)]
-	\item When $1 < d < 4$, is the test $\phi_{\spec}^{(2)}$ minimax optimal?
-	\item Are either or both of the test statistics \eqref{eqn:graph_spectral_projections_2}-\eqref{eqn:sobolev_IPM} minimax optimal over $C_d^1(\mathcal{X};R)$ for all values of $d$?
-	\item Are either or both of the test statistics \eqref{eqn:graph_spectral_projections_2}-\eqref{eqn:sobolev_IPM} minimax optimal over $W_d^{1,2}(\mathcal{X};R)$ when $d \leq 4$?
-	\item Is the test statistic \eqref{eqn:graph_spectral_projections_2}, computed over a graph with suitable choice of kernel $K$, minimax optimal over higher order derivative classes $C_d^s$ and $W^{s,2}$?
-	\item Modify the test statistic \eqref{eqn:sobolev_IPM} by replacing the function class $\Theta_{1,2}$ with
-	\begin{equation}
-	\Theta_{s,2} := \{\theta \in \Reals^n:~ C_n^{-1} \norm{B^{(s)}\theta}_2 \leq 1\} 
-	\end{equation}
-	Is the modified test statistic, computed over a graph with suitable choice of kernel $K$, minimax optimal over higher order derivative classes $C_d^s$ and $W^{s,2}$?
-	\item What is the minimax testing rate over $BV_d^{1}(\mathcal{X};R)$? Does it exhibit a phase transition analogous to the minimax estimation rate over bounded variation spaces?
-	\item Is the test statistic \eqref{eqn:total_variation_IPM} minimax optimal over $BV_d^{1}(\mathcal{X};R)$?
-	\item Is the test statistic \eqref{eqn:total_variation_IPM} minimax optimal over Sobolev and Holder function classes?
-\end{enumerate}
-
-\section{Definitions and Notation}
-
-Here we collect definitions of some common function spaces and graph operators.
-
-\subsection{Function Spaces}
-
-\paragraph{Lebesgue spaces.}
-
-We say a Borel measurable function $f: \mathcal{X} \to \Reals$ is in the space $\mathcal{L}^p(\mathcal{X})$ for $1 \leq p < \infty$ if 
-$$\norm{f}_{\mathcal{L}^p(\mathcal{X})} := \int_{\mathcal{X}} \abs{f(x)}^p \,dx < \infty$$
-and we let 
-\begin{equation*}
-\mathcal{L}^p(\mathcal{X};R) = \set{f \in \mathcal{L}^p(\mathcal{X}): \norm{f}_{\mathcal{L}^p(\mathcal{X})} < R}
-\end{equation*}
-be a ball in the Lebesgue space.
-
-\paragraph{Holder spaces.}
-
-For a given $s > 0$, let $\ell = \floor{s}$ be the largest integer strictly less than $s$. Then the $s$th Holder norm is given by
-\begin{equation*}
-\norm{f}_{C_d^{s}(\mathcal{X})} := \sum_{\abs{\alpha} < s} \norm{D^{\alpha}f}_{\infty} + \sum_{\abs{\alpha} = \ell} \sup_{x,y \in \mathcal{X}} \frac{\abs{D^{\alpha}f(y) - D^{\alpha}f(x)}}{\norm{x - y}_2^{s - \ell}}
-\end{equation*}
-and the $s$th Holder space $C_d^{s}(\mathcal{X})$ consists of all functions which are $\ell$ times continuously differentiable with finite Holder norm $\norm{\cdot}_{C_d^{s}(\mathcal{X})}$.
-Denote the Holder ball by $C_d^{s}(\mathcal{X},R) = \set{f \in C_d^{s}(\mathcal{X}): \norm{f}_{C_d^{s}(\mathcal{X})} \leq R}$.
-
-\paragraph{Sobolev spaces.}
-
-For a given $s > 0$, the Sobolev space $W_d^{s,2}(\mathcal{X})$ consists of all functions $f \in \mathcal{L}^2(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^2(\mathcal{X})$.
-The Sobolev $\{s,2\}$ norm is then
-\begin{equation*}
-\norm{f}_{W_d^{s,2}(\mathcal{X})}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^2 \,dx
-\end{equation*}
-and for a given $L > 0$, the corresponding ball is $W_d^{s,2}(\Xset; L) = \set{f: \norm{f}_{W^{s,2}(\Xset)} \leq L}$.
-
-\paragraph{Bounded Variation spaces.}
-
-For a function $f \in L^1(\mathcal{X})$ the \emph{total variation} semi-norm of $f$ is
-\begin{equation*}
-TV(f;\mathcal{X}) := \sup \left\{ \int_{\mathcal{X}} f \, \Xsetive \, \psi \,dx : \psi \in C_c^1(\mathcal{X}; \Reals^d), \abs{\psi} \leq 1 \right\};
-\end{equation*}
-and we write $BV_d(\mathcal{X})$ for the subset of functions $f \in L^1(\mathcal{X})$ which have bounded norm
-\begin{equation*}
-\norm{f}_{BV_d(\mathcal{X})} := \norm{f}_{\infty} + TV(f;\mathcal{X}).
-\end{equation*}
-For a given $R > 0$, the corresponding ball is $BV_d^{1}(\mathcal{X};R) = \set{f: \norm{f}_{BV_d(\mathcal{X})} \leq R}$.
-
-\subsection{Graph Operators}
-Let $s \geq 1$ be an integer. The $s$th-order difference operator on $G_{n,r}$, denoted $B^{(s)}$, is defined by
-\begin{equation*}
-B^{(s)} :=
-\begin{cases}
-L^{s/2},& ~~ s \textrm{ even} \\
-BL^{(s - 1)/2},& ~~ s \textrm{ odd.}
-\end{cases}
-\end{equation*}
-
-\subsection{Kernels}
-We say a kernel function $K(\cdot)$ is a $2$nd order kernel if $K$ is compactly supported on $B(0,1)$, uniformly upper bound $\abs{K(x)} \leq K_{\max} < \infty$ for all $x \in B(0,1)$, and
-\begin{equation*}
-\int K(x) \,dx = \nu_d,~~ \int x K(x) \,dx = 0.
-\end{equation*}
-Note that the uniform kernel, defined as $K(x) = \1\{\norm{x} \leq 1\}$, is a $2$nd order kernel.
-
-\subsection{Notation}
-
-We will treat $f$ interchangeably as a function $f:\Rd \to \Reals$, and as a vector $f = (f(x_1),\ldots,f(x_n))$. It should always be clear from context how we are using $f$.
-For sequences $(a_n), (b_n)$, we say $a_n = O(b_n)$ if there exists a constant $c$ such that $a_n \leq cb_n$ for all $c$. We use the notation $a \wedge b$ for the minimum of $a$ and $b$.
-
-We use the notation $\Lambda(H)$ to denote the spectrum of a matrix $H$, and $\lambda_k(H)$ to denote the $k$th smallest eigenvalue of $H$.
-
-\section{Proofs}
-
-\subsection{Proof of Theorem~\ref{thm:sobolev_testing_rate}}
-
-Our analysis will proceed according to the following steps.
-\begin{enumerate}
-	\item We upper bound the testing error of our eigenvector projection test statistic when $Y$ are viewed as random responses defined over the vertices of a fixed graph $G$.
-This upper bound will hold whenever certain functionals on the graph $G$ are themselves bounded.
-	\item We analyze the behavior of these functionals with respect to the random graph $G_{n,r}$, and bound them with high probability.
-	\item We condition on $X \subseteq \mathcal{E}$, where $\mathcal{E}$ is a high-probability set over which the relevant functionals on $G_{n,r}$ are bounded.
-We conclude that the upper bound on testing error derived in our first step holds whenever $X \subseteq \mathcal{E}$.
-\end{enumerate}
-
-\subsubsection{Step 1: Testing error on a fixed graph}
-
-Let $G = (V,E)$ be a graph over $n$ vertices $V$, and let $\beta = (\beta_1,\ldots,\beta_n) \in \Reals^n$ be a signal over the vertices.
-(Although one can think of $G$ as a realization of $G_{n,r}$ for a particular sample $X$, our theory in this section will deal with $G$ as if it were an arbitrary fixed graph.) We observe responses $Y = (y_1,\ldots,y_n)$ according to the model
-\begin{equation*}
-y_i = \beta_i + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} N(0,1)
-\end{equation*}
-Let $L$ denote the graph Laplacian of $G$, and let $0 = \lambda_1 \leq \lambda_2 \leq \ldots \leq \lambda_n$ be the ordered eigenvalues of $L$, with eigenvector $v_i$ corresponding to eigenvalue $\lambda_i$.
-Our graph spectral test statistic is
-\begin{equation*}
-T_{\spec} = \frac{1}{n}\sum_{k = 1}^{\kappa} \left(\sum_{i = 1}^{n} v_i y_i\right)^2
-\end{equation*}
-where $\kappa$ is a tuning parameter. The resulting test we will use is
-\begin{equation*}
-\phi_{\spec} = \1\{T_{\spec} \geq \frac{\kappa}{n} + t(b)\},~~\textrm{where}~ t(b) = b\sqrt{\frac{2\kappa}{n^2}}~~\textrm{for $b \geq 1$.}
-\end{equation*}
-
-In Lemma~\ref{lem:fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}$.
-Our bound on the Type II error will be stated as a function of $\beta^T L^s \beta$--a measure of the smoothness the signal $\beta$ displays over the graph $G$--as well as the $\kappa$th eigenvalue $\lambda_{\kappa}$.
-
-\begin{lemma}
-	\label{lem:fixed_graph_testing}
-	Let $1 \leq \kappa \leq n$ be an integer.
-	\begin{enumerate}
-		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}$ is upper bounded
-		\begin{equation}
-		\label{eqn:graph_spectral_type_I_error}
-		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{1}{b^2}.
-		\end{equation}
-		\item \textbf{Type II error:} For any $b$ and $\beta$ such that
-		\begin{equation}
-		\label{eqn:fixed_graph_testing_critical_radius}
-		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{\beta^T L^s \beta}{n\lambda_{\kappa}^s}
-		\end{equation}
-		the Type II error of $\phi_{\spec}$ is upper bounded,
-		\begin{equation}
-		\label{eqn:graph_spectral_type_II_error}
-		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{3}{b}.
-		\end{equation}
-	\end{enumerate}
-\end{lemma}
-
-To prove Lemma~\ref{lem:fixed_graph_testing} we will first compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}$, and then use Chebyshev's inequality to show \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.
-
-\vspace{.2 in}
-
-\textit{Mean of $T_{\mathrm{spec}}$:} Using the notation $\dotp{v}{w} = \sum_{i = 1}^{n} v_iw_i$, we have
-\begin{align*}
-\Ebb(T_{\spec}) & = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{\beta}{v_k}^2 + \Ebb\bigl( \dotp{\varepsilon}{v_k}^2 + 2 \dotp{\varepsilon}{v_k} \dotp{\beta}{v_k}\bigr)\right) \\
-& = \frac{\kappa}{n} + \frac{1}{n}\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2.
-\end{align*}
-When $\beta = 0$, this equals $\kappa/n$. Otherwise, we have the following lower bound:
-\begin{align*}
-\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2 & = \norm{\beta}_2^2 - \sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \\
-& \geq \norm{\beta}_2^2 - \frac{1}{\lambda_{\kappa}^s}\sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \lambda_k^s \\
-& \geq \norm{\beta}_2^2 - \frac{\beta^T L^s \beta}{\lambda_{\kappa}^s},
-\end{align*}
-and therefore $\Ebb(T_{\spec}) \geq \kappa/n + n^{-1}(\norm{\beta}_2^2 - \beta^T L^s \beta/\lambda_{\kappa}^s)$.
-
-\vspace{.2 in}
-
-\textit{Variance of $T_{\mathrm{spec}}$:}
-We write $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvector $v_k$ as its $k$th column.
-\begin{align*}
-\Var(T_{\spec}) & = \frac{1}{n^2} \Var(y^T V_{\kappa} V_{\kappa}^T y) \\
-& = \frac{1}{n^2} \Var((\beta + \varepsilon)^T V_{\kappa} V_{\kappa}^T (\beta + \varepsilon)) \\
-& = \frac{1}{n^2} \Var(2 \beta^T V_{\kappa} V_{\kappa}^T \varepsilon + \varepsilon^T V_{\kappa} V_{\kappa}^T \varepsilon) \\
-& \leq \frac{1}{n^2}(4 \beta^T V_{\kappa} V_{\kappa}^T \beta + 2\kappa)
-\end{align*}
-where the last inequality follows from standard properties of the Gaussian distribution.
-We now move on to showing the desired inequalities \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.
-
-\vspace{.2 in}
-
-\textit{Proof of~\eqref{eqn:graph_spectral_type_I_error}:} By Chebyshev's inequality,
-\begin{align*}
-\Pbb_{\beta = 0}\bigl(T_{\spec} \geq \frac{\kappa}{n} + t(b)\bigr)
-& \leq \Pbb_{\beta = 0}\bigl(\abs{T_{\spec} - \frac{\kappa}{n}} \geq t(b)\bigr) \\
-& \leq \frac{\Var_{\beta = 0}(T_{\spec})}{t(b)^2} = \frac{1}{b^2}.
-\end{align*}
-
-\vspace{.2 in}
-
-\textit{Proof of~\eqref{eqn:graph_spectral_type_II_error}:} For simplicity, we introduce the notation
-\begin{equation*}
-\Delta = \frac{\norm{\beta}_2^2}{n} - \frac{\beta^T L^s \beta}{n\lambda_{\kappa}^s}.
-\end{equation*}
-Assumption~\eqref{eqn:fixed_graph_testing_critical_radius} implies $\Delta \geq 2 t(b)$. Then another application of Chebyshev's inequality gives us
-\begin{align*}
-\Pbb_{\beta}\bigl(T_{\spec} \leq \frac{\kappa}{n} + t(b)\bigr) & = \Pbb_{\beta}\bigl(T_{\spec} - \Ebb_{\beta}(T_{\spec}) \leq t(b) - \Delta \bigr) \\
-& \leq \Pbb_{\beta}\bigl(\abs{T_{\spec} - \Ebb_{\beta}(T_{\spec})} \leq \Delta - t(b) \bigr) \tag{since $\Delta \geq t(b)$}	\\
-& \leq \frac{\Var_{\beta}(T_{\spec})}{(\Delta - t(b))^2} \\
-& \leq 4\frac{\Var_{\beta}(T_{\spec})}{\Delta^2} \tag{since $\Delta \geq 2t(b)$} \\
-& \leq 4\frac{2\kappa/n^2 + 4\beta^T V_{\kappa} V_{\kappa}^T \beta /n^2}{\Delta^2}.
-\end{align*}
-We handle each summand in the numerator separately. For the first term, since $\Delta \geq 2 t(b)$, we have
-\begin{equation}
-\label{eqn:spectral_type_II_error_pf1}
-\frac{2\kappa}{n^2\Delta^2} \leq \frac{1}{2b^2}.
-\end{equation}
-
-For the second term, noting that $\Delta = \beta^T V_{\kappa} V_{\kappa}^T \beta/n$, we have
-\begin{align}
-\frac{\beta^T V_{\kappa} V_{\kappa}^T \beta/n^2}{\Delta^2} & = \frac{1}{n\Delta} \nonumber \\
-& \leq \frac{1}{2nt(b)} \nonumber \\
-& = \frac{1}{2b\sqrt{2\kappa}}, \label{eqn:spectral_type_II_error_pf2}
-\end{align}
-and combining~\eqref{eqn:spectral_type_II_error_pf1} and~\eqref{eqn:spectral_type_II_error_pf2} yields~\eqref{eqn:graph_spectral_type_II_error}.
-
-\subsubsection{Step 2: Bounding neighborhood graph functionals}
-
-To make use of Lemma~\ref{lem:fixed_graph_testing} we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$.
-In particular, we will show that the following statements each hold with probability at least $1 - c/b$ for sufficiently large $n$ (Here and in what follows, $c$ denotes a constant which is fixed in $n$ and $f$ and does not depend on $b$, but may depend on other fixed quantities such as $d$, $s$, etc.):
-\begin{enumerate}[label=(E\arabic*)]
-	\item 
-	\label{event:discrete_sobolev_norm}
-	\textbf{Graph Sobolev norm:} For any $n^{-1/(2(s - 1) + d)}\leq r \leq 1$,
-	\begin{equation}
-	\label{eqn:continuous_to_discrete_sobolev_norm}
-	f^T L^s f \leq c \cdot b \cdot \norm{f}_{W^{s,2}(\Xset)}^2 \cdot n^{s + 1} r^{s(d + 2)} 
-	\end{equation}
-	\item 
-	\label{event:eigenvalue_tail_decay}
-	\textbf{Eigenvalue tail bound:} For any $\kappa = 1,\ldots,n$ and $r \leq \kappa^{-2/(d(2 +d))}$, 
-	\begin{equation}
-	\label{eqn:eigenvalue_tail_bound}
-	\lambda_{\kappa} \geq c \cdot n r^{d + 2} \kappa^{2/d}
-	\end{equation}
-	In particular, the inequality~\eqref{eqn:eigenvalue_tail_bound} is satisfied when $\kappa = n^{2d/(4s + d)}$ for any $r \leq n^{-4/((2+d)(4s + d))}$. 
-	\item 
-	\label{event:l2_norm}
-	\textbf{Empirical norm of $f$:} There exists a constant $c_1$ such that if $\norm{f}_{\Leb^2(\Xset)} \geq c_1 \cdot b \cdot n^{-ws/(4s + d)}$,
-	\begin{equation}
-	\label{eqn:l2_to_empirical_norm}
-	\norm{f}_n^2 \geq \frac{1}{b} \cdot \norm{f}_{\Leb^2}^2
-	\end{equation}
-\end{enumerate}
-
-\paragraph{Proof of~\eqref{eqn:continuous_to_discrete_sobolev_norm}:}
-
-The probabilistic bound~\eqref{eqn:continuous_to_discrete_sobolev_norm} follows from the more general Lemma~\ref{lem:roughness_functional_expectation_sobolev} by Markov's inequality. 
+\begin{proof}[Proof sketch.]
+
+	\textcolor{red}{Note}: I have actually proved Theorem~\ref{thm:higher_order_sobolev_testing_rate}, but for brevity I just included the new parts here.
+
+	From previous work, we have that for any $X \in \Xset^n$ such that
+	\begin{equation*}
+	\norm{f}_n^2 \geq 2b\sqrt{\frac{\kappa}{n^2}} + \frac{f^T L^s f}{n(nr^{(d + 2)}\kappa^{1/d})^s}
+	\end{equation*}
+	the type II error of the test $\phi_{\spec}$ conditional on $X$ is upper bounded by $c/b$. By some standard calculations, it is therefore sufficient to show that
+	\begin{equation}
+	\label{eqn:higher_order_sobolev_testing_rate_pf1}
+	\norm{f}_n^2 \geq \frac{c}{b}\norm{f}_{\Leb^2(\Xset)}
+	\end{equation}
+	and
+	\begin{equation}
+	\label{eqn:higher_order_sobolev_testing_rate_pf2}
+	f^T L^s f \leq b \cdot n^{s + 1} r^{d(s + 2)} \norm{f}_{W_d^{s,2}(\Xset)}
+	\end{equation}
+	are each satisfied with probability at least $1 - c/b$. 
+\end{proof}
+
+\section{Proof of~\eqref{eqn:higher_order_sobolev_testing_rate_pf1}}
+
+We find it more convenient to work with a normalized version of the graph Sobolev seminorm, 
+\begin{equation*}
+R_{s,n}(f) = \frac{1}{n^{s + 1}r^{d(s + 2)}} f^T L^s f. 
+\end{equation*}
+We have that in expectation, the roughness functional $R_{s,n}(f)$ is (up to constants), no greater than the Sobolev norm $\norm{f}_{W_d^{1,2}(\mathcal{X})}$.
 \begin{lemma}
 	\label{lem:roughness_functional_expectation_sobolev}
-	Let $\Xset$ be a Lipschitz domain, and let $s \geq 1$ be an integer.
+	Let $\Xset$ be a Lipschitz domain.
 Suppose that $f \in W^{s,2}(\Xset)$, and further that $p \in C^{s-1}(\Xset;p_{\max})$ for some constant $p_{\max}$.
-Then for any $2$nd-order kernel $K$ and any $n^{-1/(2(s - 1) + d)} \leq r(n) \leq 1$, the expected graph Sobolev seminorm is upper bounded
+Then for any $2$nd-order kernel $K$ and any $n^{-1/(2(s - 1) + d)} \leq r(n) \leq 1$, for sufficiently large $n$ the expected graph Sobolev seminorm is upper bounded
 	\begin{equation}
 	\label{eqn:roughness_functional_expectation_sobolev}
-	\Ebb\bigl[f^T L^s f\bigr] \leq c \cdot \norm{f}_{W_d^{s,2}(\Xset)}^2 \cdot n^{s + 1}r^{s(d + 2)}
-	\end{equation}
-\end{lemma}
-
-The proof of Lemma~\ref{lem:roughness_functional_expectation_sobolev} is lengthy, and we defer it to Section~\ref{sec:technical_lemma_proofs}.
-
-\paragraph{Proof of~\eqref{eqn:eigenvalue_tail_bound}:}
-
-We prove~\eqref{eqn:eigenvalue_tail_bound} by comparing $G_{n,r}$ to the tensor product of a $d$-dimensional lattice and a complete graph.
-The latter is a highly structured graph with known eigenvalues, which as we will see are sufficiently lower bounded for our purposes.
-
-Let $\wt{r} = r/(3(\sqrt{d} + 1)), M = (1/\wt{r})^d, N = n\wt{r}^d$. Assume without loss of generality that $M$ and $N$ are integers. Additionally, for $t = n^{1/d}$ and $m = M^{1/d}$ let 
-\begin{equation*}
-\overline{X} = \set{\frac{1}{t}(k_1,\ldots,k_d): k \in [t]^d},~~ \overline{Z} = \set{\frac{1}{m}(j_1,\ldots,j_d): j \in [m]^d}.
-\end{equation*}
-For a given $\overline{z}_j \in \overline{Z}$, we write $Q(z_j) = m^{-1}[j_1 - 1,j_1] \times \cdots \times m^{-1}[j_d - 1,j_d]$ for the cube of side length $1/m$ with $z_j$ at one corner.
-
-Consider the graph $H = (\overline{X}, E_H)$, where $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$ if
-\begin{equation*}
-\textrm{there exists}~\ol{z}_i, \ol{z}_j \in \ol{Z}~\textrm{such that}~\ol{x}_k \in Q(\ol{z}_i),~ \ol{x}_\ell \in Q(\ol{z}_j),~\textrm{and}~\norm{i - j}_1 \leq 1.
-\end{equation*}
-On the one hand $H \cong \ol{G}^M_d \otimes K_N$ where $\ol{G}^M_d$ is the $d$-dimensional lattice on $M$ nodes, and $K_N$ is the complete graph on $N$ nodes.
-On the other hand, we now show that with high probability $G_{n,r} \succeq H$.
-If $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$, then there exist $\ol{z}_i, \ol{z}_j$ such that
-\begin{equation*}
-\norm{\ol{x}_k - \ol{x}_{\ell}}_2 \leq m^{-1} + \norm{\ol{x}_k - \ol{z}_{i}}_2 + \norm{\ol{x}_{\ell} - \ol{z}_{j}}_2 \leq \wt{r}(1 + \sqrt{d}) = r/3.
-\end{equation*}
-
-Assuming~\eqref{eqn:slepcev_transport_distance} holds, if $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$, then for sufficiently large $n$
-\begin{equation*}
-\norm{\pi(\ol{x}_k) - \pi(\ol{x}_\ell)}_2 \leq 2 c \left(\frac{\log n}{n}\right)^{1/d} + \frac{r}{3} \leq r,
-\end{equation*}
-implying that $(\pi(\ol{x}_k), \pi(\ol{x}_{\ell})) \in E$. Therefore, $G_{n,r} \succeq \ol{G}^M_d \otimes K_N$ whenever~\eqref{eqn:slepcev_transport_distance} holds.
-
-The eigenvalues of lattices and complete graphs are known to satisfy, respectively
-\begin{equation*}
-\lambda_k(\ol{G}^{M}_d) \geq \frac{k^{2/d}}{M^{2/d}}~\textrm{for $k = 0,\ldots,M - 1$},~~ \textrm{and}~\lambda_{j}(K_N) \geq N\1\{j > 0\}~\textrm{for $j = 0,\ldots,N-1$.}
-\end{equation*}
-and by standard facts regarding the eigenvalues of tensor product graphs, we have that the spectrum $\Lambda(H)$ satisfies
-\begin{equation*}
-\Lambda(H) = \set{N\lambda_k(\ol{G}^{M}_d) + M\lambda_j(K_N): \textrm{for $k = 0,\ldots,M - 1$ and $j = 0,\ldots,N-1$}}
-\end{equation*}
-For all $j = 1,\ldots,N-1$, we have that $M\lambda_j(K_N) = MN = n$. Therefore,
-\begin{align*}
-\lambda_{\kappa}(H) & \geq \{n \wedge N\lambda_{\kappa}(\ol{G}^{M}_d)\} \\
-& \geq \{n \wedge n\wt{r}^d\frac{\kappa^{2/d}}{M^{2/d}}\} \\
-& \geq \{n \wedge (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d}\} \\
-& \geq (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d},
-\end{align*}
-where the last inequality is satisfies since $r \leq \kappa^{-2/(d(d + 2))}$.
-
-Finally, Theorem~\ref{thm:slepcev_transport_distance} establishes that \eqref{eqn:slepcev_transport_distance} occurs with probability at least $1 - c/n$.
-Since $\lambda_{\kappa}(G_{n,r}) \geq \lambda_{\kappa}(H)$ when~\eqref{eqn:slepcev_transport_distance} is satisfied, this completes the proof of~\eqref{eqn:eigenvalue_tail_bound}.
-
-\paragraph{Proof of~\eqref{eqn:l2_to_empirical_norm}:}
-
-\begin{lemma}
-	\label{lem:empirical_norm_sobolev}
-	Let $\Xset$ be a Lipschitz domain over which the density is upper and lower bounded 
-	\begin{equation*}
-	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty ~~\textrm{for all $x \in \Xset$,}
-	\end{equation*}
-	and let $f \in W^{s,2}(\Xset)$.Then for any $b \geq 1$, there exists $c_1$ such that if 
-	\begin{equation}
-	\norm{f}_{\Leb^2(\Xset)} \geq 
-	\begin{cases*}
-	c_1 \cdot b \cdot \norm{f}_{W^{s,2}(\Xset)} \cdot \max\{n^{-1/2},n^{-s/d}\},~~\textrm{if}~2s \neq d \\
-	c_1 \cdot b \cdot \norm{f}_{W^{s,2}(\Xset)} \cdot n^{-a/2},~~\textrm{if}~ 2s = d ~\textrm{for any}~ 0 < a < 1
-	\end{cases*}
-	\end{equation}
-	then,
-	\begin{equation}
-	\label{eqn:l2_to_empirical_norm_sobolev}
-	\mathbb{P}\left[\norm{f}_n^2 \geq \frac{1}{b} \Ebb\bigl[\norm{f}_n^2\bigr]\right] \geq 1 - \frac{5}{b}
-	\end{equation}
-	where $c_1$ and $c_2$ are constants which may depend only on $s$, $\Xset$, $d$, $p_{\min}$ and $p_{\max}$.
-\end{lemma}
-The lower bound~\eqref{eqn:l2_to_empirical_norm} results from the more general Lemma~\ref{lem:empirical_norm_sobolev}, which can be verified by checking the various orderings of $2s/(4s + d)$, $s/d$ and $1/2$ whenever $4s < d$.
-\begin{proof}[Proof of Lemma~\ref{lem:empirical_norm_sobolev}]
-	To prove~\eqref{eqn:l2_to_empirical_norm_sobolev} we will show
-	\begin{equation*}
-	\mathbb{E}\bigl[\norm{f}_n^4\bigr] \leq \left(1 + \frac{1}{b^2}\right)\cdot \left(\mathbb{E}\bigl[\norm{f}_n^2\bigr]\right)^2
-	\end{equation*}
-	whence the claim follows from the Paley-Zygmund inequality (Lemma~\ref{lem:paley_zygmund}).
-Since $p \leq p_{\max}$ is uniformly bounded, we can relate $\mathbb{E}\bigl[\norm{f}_n^4\bigr]$ to the $\Leb^4$ norm,
-	\begin{equation*}
-	\mathbb{E}\bigl[\norm{f}_n^4\bigr] = \frac{(n-1)}{n}\left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + \frac{\mathbb{E}\Bigl[\bigl(f(x_1)\bigr)^4\Bigr]}{n} \leq \left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + p_{\max}^2\frac{\norm{f}_{\Leb^4}^4}{n}.
-	\end{equation*}
-	We will use a Sobolev inequality to relate $\norm{f}_{\Leb^4}$ to $\norm{f}_{W_d^{s,2}(\Xset)}$.
-The nature of this inequality depends on the relationship between $s$ and $d$ (see Theorem 6 in Section 5.6.3 of \textcolor{red}{Evans} for a formal statement), so from this point on we divide our analysis into three cases: (i) the case where $2s > d$, (ii) the case where $2s < d$, and (iii) the borderline case $2s = d$.
-
-	\paragraph{Case 1: $2s > d$.}
-	When $2s > d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in C^{\gamma}(\overline{\Xset})$ for some $\gamma > 0$ which depends on $s$ and $d$, with the accompanying estimate
-	\begin{equation*}
-	\sup_{x \in \Xset} \abs{f(x)} \leq \norm{f}_{C^{\gamma}(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}.
-	\end{equation*}
-	Therefore,
-	\begin{align*}
-	\norm{f}_{\Leb^4}^4 & = \int_{\Xset} [f(x)]^4 \,dx \\
-	& \leq \left(\sup_{x \in \Xset} \abs{f(x)}\right)^2 \cdot \int_{\Xset} [f(x)]^2 \,dx \\
-	& \leq c \norm{f}_{W^{s,2}(\Xset)}^2 \cdot \norm{f}_{\Leb^2(\Xset)}^2.
-	\end{align*}
-	Since by assumption
-	\begin{equation*}
-	\norm{f}_{\Leb^2(\Xset)}^2 \geq c_1^2 \cdot b^2 \cdot \norm{f}_{W_d^{s,2}(\Xset)}^2 \cdot \frac{1}{n},
-	\end{equation*}
-	we have
-	\begin{equation*}
-	p_{\max}^2\frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c \norm{f}_{W^{s,2}(\Xset)}^2 \cdot \frac{\norm{f}_{\Leb^2(\Xset)}^4}{n \norm{f}_{\Leb^2(\Xset)}^2} \leq c \frac{\norm{f}_{\Leb^2(\Xset)}^4}{c_1^2 b^2} \leq \frac{\Ebb\bigl[\norm{f}_n^2\bigr]}{b^2},
-	\end{equation*}
-	where the last inequality follows by taking $c_1$ sufficiently large.
-
-	\paragraph{Case 2: $2s < d$.}
-	When $2s < d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in \Leb^q(\Xset)$ for $q = 2d/(d - 2s)$, and moreover that
-	\begin{equation*}
-	\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}.
-	\end{equation*}
-	Since $4 = 2\theta + (1 - \theta)q$ for $\theta = 2 - d/(2s)$, Lyapunov's inequality implies
-	\begin{equation*}
-	\norm{f}_{\Leb^4(\Xset)}^4 \leq \norm{f}_{\Leb^2}^{2\theta} \cdot \norm{f}_{\Leb^q(\Xset)}^{(1 - \theta)q} \leq c \norm{f}_{\Leb^2(\Xset)}^{4} \cdot \left(\frac{\norm{f}_{W^{s,2}(\Xset)}}{\norm{f}_{\Leb^2(\Xset)}}\right)^{d/s}.
-	\end{equation*}
-	By assumption, $\norm{f}_{\Leb^2(\Xset)} \geq c_1 b \norm{f}_{W^{s,2}(\Xset)} n^{-s/d}$, and therefore
-	\begin{equation*}
-	p_{\max}^2 \frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c\norm{f}_{\Leb^2(\Xset)}^4 \left(\frac{\norm{f}_{W^{s,2}(\Xset)}}{n^{s/d}\norm{f}_{\Leb^2(\Xset)}}\right)^{d/s} \leq \frac{c\norm{f}_{\Leb^2(\Xset)}^4}{c_1b^{d/s}} \leq \frac{\norm{f}_{\Leb^2(\Xset)}^4}{b^2}.
-	\end{equation*}
-	where the last inequality follows when $c_1$ is sufficiently large, and keeping in mind that $d/s > 2$ and $b \geq 1$.
-
-	\paragraph{Case 3: $2s = d$.}
-	Assume $f$ satisfies~\eqref{eqn:paley_zygmund_1} for a given $0 < a < 1$.
-When $2s = d$, since $\Xset$ is a Lipschitz domain we have that $f \in L^q(\Xset)$ for any $q < \infty$, with the accompanying estimate
-	\begin{equation*}
-	\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}.
-	\end{equation*}
-	In particular the above holds for $q = 2/(1 - a)$ when $1/2 < a < 1$, and for any $q > 4$ when $0 < a < 1/2$. Using Lyapunov's inequality as in the previous case then implies the desired result.
-\end{proof}
-
-\subsubsection{Step 3: Putting the pieces together}
-
-We note that for all possible values of $X \in \Xset^n$, under the null hypothesis $f = f_0 = 0$ and therefore $\beta = (f(x_1),\ldots,f(x_n)) = 0$ as well.
-Therefore by~\eqref{eqn:graph_spectral_type_I_error}, we have the following bound on Type I error:
-\begin{equation}
-\Ebb_{f_0}(\phi_{\mathrm{spec}}) = \mathbb{E}(\mathbb{E}_{\beta = 0}(\phi_{\spec}) | X) \leq \frac{1}{b^2}.
-\end{equation}
-
-Now, we bound Type II error under the assumption $f \in W^{s,2}(\mathcal{X};R), p \in C^{s - 1}(\Xset;p_{\max})$ uniformly bounded away from $0$ and $\infty$ over $\Xset$, and 
-\begin{equation}
-\label{eqn:critical_radius_1}
-\norm{f}_{\Leb^2}^2 \geq \epsilon^2 = c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4s + d)}\} \cdot n^{-4s/(4s + d)}.
-\end{equation}
-Choosing $n^{-1/(2(s - 1) + d)}\leq r(n) \leq n^{-4/((2+d)(4s + d))}$, we may therefore apply our conclusions in Step 2; namely, that for every possible choice of $f \in W^{s,2}(\Xset;R)$ there exists a good set $\mathcal{E}_f \subseteq \Xset^n$ with $\Pbb(\mathcal{E}_f) \geq 1 - c/b$ such that each of \eqref{eqn:continuous_to_discrete_sobolev_norm}, \eqref{eqn:eigenvalue_tail_bound}, and \eqref{eqn:l2_to_empirical_norm_sobolev} hold for all $X \subseteq \mathcal{E}_f$.
-The choice $\kappa = (nR^2)^{2d/(4s + d)}$ balances the squared bias and variance terms on the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius}, and we have that for all $X \subseteq \mathcal{E}_f$
-\begin{align*}
-2b\sqrt{\frac{2\kappa}{n^2}} + \frac{f^T L^s f}{n\lambda_{\kappa}^s} & \leq 2bR^{2d/(4s + d)}n^{-4s/(4s+d)} + cbR^2\frac{n^{s} r^{s(d + 2)}}{\lambda_{\kappa}^s} \tag{by \eqref{eqn:continuous_to_discrete_sobolev_norm}} \\
-& \leq cbR^{2d/(4s + d)}n^{-4s/(4s+d)} \tag{by \eqref{eqn:eigenvalue_tail_bound}} \\
-& \leq \frac{1}{b}\norm{f}_{\Leb^2(\Xset)}^2 \\
-& \leq \frac{1}{n}\sum_{i = 1} \bigl[f(x_i)\bigr]^2. \tag{by \eqref{eqn:l2_to_empirical_norm_sobolev}}
-\end{align*}
-where the last two inequalities follow for a suitably large choice of $c_1$ in~\eqref{eqn:critical_radius_1}.
-We conclude that for all $X \subseteq \mathcal{E}_f$, the inequality \eqref{eqn:fixed_graph_testing_critical_radius} is satisfied with respect to $\beta = (f(x_1),\ldots,f(x_n))$ and $G = G_{n,r}$.
-As a result the worst-case Type II error is bounded
-\begin{equation*}
-\sup_{\substack{f \in W^{s,2}(\Xset;R), \\ \norm{f}_{\Leb^2(\Xset)} \geq \epsilon}}\mathbb{E}_{f}(1 - \phi_{\spec}) \leq \sup_{\substack{f \in W^{s,2}(\Xset;R), \\ \norm{f}_{\Leb^2(\Xset)} \geq \epsilon}} \mathbb{E}\bigl[\mathbb{E}_{\beta}(1 - \phi_{\spec}|X \in \mathcal{E}_f)\bigr] + \frac{c}{b} \leq \frac{3 + c}{b},
-\end{equation*}
-completing the proof of Theorem~\ref{thm:sobolev_testing_rate} upon proper choice of constants $c_1,c_2$ in that Theorem.
-
-\clearpage
-
-\subsection{Proof of Proposition~\ref{prop:L4_testing_rate}}
-
-Let $T_{\mathrm{mean}} = \frac{1}{n}\sum_{i = 1}^{n} y_i^2$. The expectation of $T_{\mathrm{mean}}$ is
-\begin{equation*}
-\Ebb(T_{\mathrm{mean}}) = \mathbb{E}(f^2(x_1)) + 1,
-\end{equation*}
-and the variance can be upper bounded
-\begin{equation*}
-\Var(T_{\mathrm{mean}}) \leq \frac{1}{n}(3 + p_{\max} \norm{f}_{\Leb^4}^4 + p_{\max}\norm{f}_{\Leb^2}^2) = \frac{c}{n}.
-\end{equation*}
-Since $\mathbb{E}(f^2(x_1)) \geq p_{\min} \epsilon^2$, when $\epsilon^2 \gtrsim b\cdot n^{-1/2}$ we can apply Chebyshev's inequality to obtain the claimed result.
-
-\subsection{Proof of Theorem~\ref{thm:twosample_sobolev_testing_rate}}
-
-As mentioned previously, to prove Theorem~\ref{thm:twosample_sobolev_testing_rate}, we relate the two sample model to the following regression model: we observe
-$X = \{x_1,\ldots,x_n\} \sim \mu$ (where we recall $\mu = (P + Q)/2$), and associated labels
-\begin{equation*}
-a_i = 
-\begin{cases}
-1, & \textrm{with probability}~ \frac{p(x_i)}{p(x_i) + q(x_i)} \\
--1, & \textrm{with probability}~ \frac{q(x_i)}{p(x_i) + q(x_i)}
-\end{cases}
-\end{equation*}
-where $a_i$ is conditionally independent of $x_j,a_j$ given $x_i$.
-
-\textcolor{red}{TODO:} Complete the above argument.
-
-Our analysis will trace a similar path to the proof of Theorem~\ref{thm:sobolev_testing_rate}, proceeding according to the following steps:
-\begin{enumerate}
-	\item We upper bound the testing error of our eigenvector projection test statistic when $a$ are viewed as random responses defined over the vertices of a fixed graph $G$.
-This upper bound will hold whenever certain functionals on the graph $G$ are themselves bounded.
-One of these functionals will be a measure of eigenvector incoherence.
-	\item We analyze the behavior of these functionals with respect to the random graph $G_{n,r}$, and bound them with high probability.
-	\item We condition on $X \subseteq \mathcal{E}$, where $\mathcal{E}$ is a high-probability set over which the relevant functionals on $G_{n,r}$ are bounded.
-We conclude that the upper bound on testing error derived in our first step holds whenever $X \subseteq \mathcal{E}$.
-\end{enumerate}
-
-\subsubsection{Step 1: Testing error on a fixed graph}
-
-Let $G = (V,E)$ be a graph over vertices $V = \set{v_1,\ldots,v_n}$, and let $\beta^{(p)} = (\beta_1^{(p)},\ldots,\beta_n^{(p)}) \in \Reals^n$, $\beta^{(q)} = (\beta_1^{(q)},\ldots,\beta_n^{(q)}) \in \Reals^n$ be non-negative signals over the vertices $V$.
-We observe labels $a = (a_1,\ldots,a_n)$ according to the model
-\begin{equation*}
-a_i = 
-\begin{cases*}
-1, & \textrm{with probability $\frac{\betap_i}{\betap_i + \betaq_i}$} \\
--1, & \textrm{with probability $\frac{\betaq_i}{\betap_i + \betaq_i}$}
-\end{cases*}
-\end{equation*}
-Letting $L = VSV^T$ be the spectral decomposition of the Laplacian $L$ of $G$, our graph spectral test statistic is 
-\begin{equation*}
-T_{\spec}^{(2)} = \frac{1}{n}\sum_{k = 1}^{\kappa} \left(\sum_{i = 1}^{n} v_i a_i\right)^2
-\end{equation*}
-where $\kappa$ is a tuning parameter. The resulting test we will use is
-\begin{equation*}
-\phi_{\spec}^{(2)} = \1\{T_{\spec}^{(2)} \geq \frac{\kappa}{n} + t(b)\},~~\textrm{where}~ t(b) = b\sqrt{\frac{2\kappa}{n^2}}~~\textrm{for $b \geq 1$.}
-\end{equation*}
-
-Let $\Pi_{\max}(\kappa;G)$ be a measure of the incoherence of the eigenvectors $v_1,\ldots,v_\kappa$, given by
-\begin{equation*}
-\Pi_{\max}(\kappa;G) := \max_{i = 1,\ldots,n} \left\{\sum_{k = 1}^{\kappa} v_{k,i}^2\right\}
-\end{equation*}
-In Lemma~\ref{lem:twosample_fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}^{(2)}$.
-Our bound on the Type II error will be stated as a function of $\Pi_{\max}(\kappa;G)$ as well as the $\kappa$th eigenvalue $s_{\kappa}$ and the smoothness functional $S_2(\beta;G)$.
-We use the notation $\varDelta^{(p,q)} = (\betap - \betaq)/(\betap + \betaq)$.
-
-\begin{lemma}
-	\label{lem:twosample_fixed_graph_testing}
-	Let $1 \leq \kappa \leq n$ be an integer.
-	\begin{enumerate}
-		\item \textbf{Type I error:} Under the null hypothesis $\betap = \betaq$, the Type I error of $\phi_{\spec}^{(2)}$ is upper bounded
-		\begin{equation}
-		\label{eqn:twosample_graph_spectral_type_I_error}
-		\mathbb{E}_{\betap,\betap}(\phi_{\spec}^{(2)}) \leq \frac{1}{b^2}.
-		\end{equation}
-		\item \textbf{Type II error:} Suppose $\Pi_{\max}(\kappa;G) \leq 1$. Then for any $b$, $\betap$ and $\betaq$ such that
-		\begin{equation}
-		\label{eqn:twosample_fixed_graph_testing_critical_radius}
-		\frac{1}{n} \sum_{i = 1}^{n} (\varDelta_i^{(p,q)})^2 \geq \frac{1}{1 - \Pi_{\max}(\kappa;G)}\left(2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_2(\varDelta;G)}{ns_{\kappa}}\right)
-		\end{equation}
-		the Type II error of $\phi_{\spec}^{(2)}$ is upper bounded,
-		\begin{equation}
-		\label{eqn:twosample_graph_spectral_type_II_error}
-		\mathbb{E}_{\betap,\betaq}(1 - \phi_{\spec}^{(2)}) \leq \frac{3}{b}.
-		\end{equation}
-	\end{enumerate}
-\end{lemma}
-
-\paragraph{Proof of Lemma~\ref{lem:twosample_fixed_graph_testing}.}
-
-To prove Lemma~\ref{lem:twosample_fixed_graph_testing} we will compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}^{(2)}$.
-The inequalities \eqref{eqn:twosample_graph_spectral_type_I_error} and \eqref{eqn:twosample_graph_spectral_type_II_error} can then be computed using Chebyshev's inequality in a manner very similar to the proof of Lemma~\ref{lem:fixed_graph_testing}, and we omit the details.
-Let $w_i := a_i - \vardeltapq_{i}$, and note that $\Ebb(w_i) = 0$ and
-\begin{equation*}
-\Ebb(w_i^2) = \frac{\betap_i \betaq_i}{(\betap_i + \betaq_i)^2} = 1 - (\vardeltapq_{i})^2,~~ \Ebb(w_i^3) = \frac{1}{2}(1 - (\vardeltapq_{i})^2)\vardeltapq_{i}
-\end{equation*}
-
-\vspace{.2 in}
-
-\textit{Mean of $T_{\spec}^{(2)}$}:
-
-We have
-\begin{align*}
-\Ebb(T_{\spec}^{(2)}) & = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{\vardeltapq}{v_k}^2 + \Ebb\bigl(\dotp{w}{v_k}^2 + 2 \dotp{w}{v_k} \dotp{\vardeltapq}{v_k}\bigr)\right) \\
-& = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{1 - (\vardeltapq)^2}{v_k^2} + \dotp{\vardeltapq}{v_k}\right) \\
-& \geq \frac{\kappa}{n} + \left(\sum_{k = 1}^{\kappa}\dotp{\vardeltapq}{v_k} - \dotp{(\vardeltapq)^2}{v_k^2}\right)
-\end{align*}
-When $\vardeltapq = 0$, this equals $\kappa/n$. Otherwise, we have the following upper bound:
-\begin{align*}
-\sum_{k = 1}^{\kappa} \dotp{(\vardeltapq)^2}{v_k^2}  & = \sum_{i = 1}^{n} (\vardeltapq_{i})^2 \left(\sum_{k = 1}^{\kappa} v_{k,i}^2 \right)\\
-& \leq \Pi_{\max}(\kappa,G) \cdot \sum_{i = 1}^{n} (\vardeltapq_{i})^2 = \Pi_{\max}(\kappa,G) \cdot \norm{\vardeltapq}_2^2.
-\end{align*}
-Additionally, $\sum_{k = 1}^{\kappa}\dotp{\vardeltapq}{v_k} \geq \norm{\vardeltapq}_2^2 - S_2(\beta;G)/s_{\kappa}$ by reasoning given in the proof of Lemma~\ref{lem:fixed_graph_testing}, and as a result
-\begin{equation*}
-\Ebb(T_{\spec}^{(2)}) \geq \frac{\kappa}{n} -  \frac{\bigl(1 - \Pi_{\max}(\kappa;G)\bigr)}{n} \norm{\vardeltapq}_2^2 - \frac{S_2(\beta;G)}{ns_{\kappa}}
-\end{equation*}
-
-\vspace{.2 in}
-\textit{Variance of $T_{\spec}^{(2)}$:}
-
-We have
-\begin{align}
-\Var(T_{\spec}^{(2)}) & = \frac{1}{n^2}\Var(2\dotp{V_{\kappa}V_{\kappa}^T\vardeltapq}{w} + \dotp{V_{\kappa}V_{\kappa}^Tw}{w}) \nonumber \\
-& = \frac{1}{n^2}\Bigl\{4\underbrace{\Var(\dotp{V_{\kappa}V_{\kappa}^T\vardeltapq}{w})}_{=: V_1} + 2\underbrace{\Cov(\dotp{V_{\kappa}V_{\kappa}^T\vardeltapq}{w},\dotp{V_{\kappa}V_{\kappa}^Tw}{w})}_{=: K_1} + \underbrace{\Var(\dotp{V_{\kappa}V_{\kappa}^Tw}{w})}_{=:V_2}\Bigr\}, \label{eqn:var}
-\end{align}
-and we now upper bound each of the three terms on the right hand side of the previous display.
-
-\textbf{Upper bound on $V_1$:}
-Let $\Sigma := \Cov(w)$ be the covariance matrix of $w$. Noting that $\Sigma \preceq I$, we have
-\begin{equation}
-\label{eqn:var1}
-\Var(\dotp{V_{\kappa}V_{\kappa}^T\vardeltapq}{w}) =(\vardeltapq)^T V_{\kappa}V_{\kappa}^T \Sigma V_{\kappa}V_{\kappa}^T \vardeltapq \leq \norm{V_{\kappa}V_{\kappa}^T\vardeltapq}^2.
-\end{equation}
-
-\textbf{Upper bound on $K_1$:}
-Noting that $\Ebb(\dotp{V_{\kappa}V_{\kappa}^T \vardeltapq}{w}) = 0$, we have that
-\begin{align}
-K_1 & = \Ebb\left[\dotp{V_{\kappa}V_{\kappa}^T \vardeltapq}{w} \dotp{V_{\kappa}V_{\kappa}^T w}{w}\right] \nonumber \\
-& = \sum_{i = 1}^{n} \sum_{j = 1}^{n} \sum_{i' = 1}^{n} \Ebb\left[w_i w_j (V_{\kappa}V_{\kappa}^T)_{ij} w_{i'} (V_{\kappa}V_{\kappa}^T\vardeltapq)_i\right] \nonumber \\
-& = \sum_{i = 1}^{n} \Ebb\left[w_i^3\right] (V_{\kappa}V_{\kappa}^T)_{ii} (V_{\kappa}V_{\kappa}^T \vardeltapq)_i \nonumber \\
-& = \frac{1}{2}\sum_{i = 1}^{n} (1 - (\vardeltapq_{i})^2)(\vardeltapq_{i}) (V_{\kappa} V_{\kappa}^T)_{ii}  (V_{\kappa}V_{\kappa}^T\vardeltapq)_i \nonumber \\
-& \leq \norm{V_{\kappa}V_{\kappa}^T \vardeltapq} \cdot \left(\sum_{i = 1}^{n} (\vardeltapq_{i})^2 (V_{\kappa} V_{\kappa}^T)_{ii}^2\right)^{1/2} \nonumber \\
-& \leq \Pi(\kappa,G) \cdot \norm{V_{\kappa}V_{\kappa}^T \vardeltapq} \cdot \norm{\vardeltapq}  \label{eqn:var2} 
-\end{align}
-
-\textbf{Upper bound on $V_2$:}
-$V_2$ is a variance of a sum, which we re-express as the sum of covariances:
-\begin{align*}
-V_2 & = \Var(\dotp{V_{\kappa}V_{\kappa}^T w}{w}) \\
-& = \sum_{i,j,i',j' = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ij} (V_{\kappa}V_{\kappa}^T)_{i'j'} \Cov(w_i w_j, w_{i'} w_{j'}).
-\end{align*}
-This covariance will be non-zero only when $i = i' \neq j = j'$, $i = j' \neq j = i'$, or $i = i' = j = j'$, and therefore
-\begin{align}
-V_2 & = 2\sum_{i = 1}^{n}\sum_{j = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ij}^2 \Var(w_iw_j) + \sum_{i = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ii}^2 \Var(w_i^2) \nonumber\\
-& \leq 2\sum_{i = 1}^{n}\sum_{j = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ij}^2 + \sum_{i = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ii}^2 \leq 3 \mathrm{tr}\bigl((V_{\kappa}V_{\kappa}^T)^2\bigr) = 3\kappa.
-\label{eqn:var3}
-\end{align}
-
-Now, plugging \eqref{eqn:var1}, \eqref{eqn:var2}, and \eqref{eqn:var3} back into \eqref{eqn:var}, we obtain
-\begin{equation}
-\label{eqn:var_5}
-\Var(T_{\spec}^{(2)}) \leq \frac{1}{n^2}\left\{4\norm{V_{\kappa}V_{\kappa}^T \vardeltapq}^2 + \Pi(\kappa;G) \cdot \norm{V_{\kappa}V_{\kappa}^T \vardeltapq} \cdot \norm{\vardeltapq} + 3\kappa \right\}.
-\end{equation}
-
-\subsubsection{Step 2: Bounding neighborhood graph functionals}
-
-Note that since $p,q \in W^{1,2}(\mathcal{X};R)$, we have $f := p - q \in W^{1,2}(\mathcal{X};2R)$.
-Therefore the bounds~\eqref{eqn:continuous_to_discrete_sobolev_norm} and \eqref{eqn:l2_to_empirical_norm} apply to $S_2(\varDelta,G_{n,r})$ and $(1/n)\sum_{i = 1}^{n}\varDelta_i^2$, respectively.
-The bound~\eqref{eqn:eigenvalue_tail_bound} continues to apply to $s_{\kappa}$.
-What remains is to upper bound the incoherence parameter
-\begin{equation}
-\label{eqn:pi_max}
-\Pi_{\max}(\kappa,G_{n,r}) \leq \frac{1}{2}~~\textrm{when}~ \kappa = n^{2/5}~\textrm{and}~ r = \log^a n \cdot \left(\frac{\log n}{n}\right)
-\end{equation}
-with probability $1 - o(1)$ as $n \to \infty$.
-
-Our proof of~\eqref{eqn:pi_max} will proceed by relating $G_{n,r}$ to a graph which satsifies a strict notion of incoherence, the chain graph.
-When $G = \overline{G}$, the incoherence parameter $\Pi_{\max}$ is upper bounded
-\begin{equation}
-\Pi_{\max}(\kappa,\overline{G}) \leq \frac{\kappa}{n}.
-\end{equation}
-To prove that the weaker bound \eqref{eqn:pi_max} holds with respect to $G_{n,r}$, we will show that the two graphs $\overline{G}$ and $G_{n,r}$ are $\delta$-spectrally similar, meaning
-\begin{equation}
-\label{eqn:spectral_similarity}
-(1 - \delta_n) x^T L x \leq x^T \overline{L} x \leq (1 + \delta_n) x^T L x~~\textrm{for all}~x \in \Reals^n, \delta_n = c \cdot (\log n)^{h(a,1) + da}\cdot \frac{\log n}{n}
-\end{equation}
-with probability $1 - o(1)$ as $n \to \infty$.
-The proof of~\eqref{eqn:spectral_similarity} relies heavily on bounding the transport distance between $X$ and $\overline{X}$, i.e.
-\eqref{eqn:slepcev_transport_distance}.
-Then the following technical Lemma along with \eqref{eqn:spectral_similarity} will imply \eqref{eqn:pi_max}.
-\begin{lemma}
-	\label{lem:pi_max_pf_1}
-	Let $G$ satisfy~\eqref{eqn:spectral_similarity} for a given $0 \leq \delta < 1$. Then for any $R \gtrsim n \delta^{1/2} (s_{n}^{1/2} \vee 1)$ and $k \leq n/8$,
-	\begin{equation*}
-	v_{k,i}^2 \lesssim \frac{1}{n}\left(R + \frac{n^4 \delta^2 s_{n}^2}{R^3}\right) \quad \textrm{for every $i = 1,\ldots,n$.}
-	\end{equation*} 
-\end{lemma}
-
-\textcolor{red}{TODO}: Explain how Lemma~\ref{lem:pi_max_pf_1} implies~\eqref{eqn:pi_max}.
-
-At a high level, Lemma~\ref{lem:pi_max_pf_1} is proved through repeated applications of the Davis-Kahan Theorem.
-The full proof is long and we delay presenting it until Section~\ref{sec:technical_lemma_proofs}.
-
-\subsubsection{Step 3: Conclusion}
-
-With Lemma~\ref{lem:twosample_fixed_graph_testing} as well as \eqref{eqn:continuous_to_discrete_sobolev_norm}, \eqref{eqn:eigenvalue_tail_bound}, \eqref{eqn:l2_to_empirical_norm} and \eqref{eqn:pi_max} in hand, the proof of Theorem~\ref{thm:twosample_sobolev_testing_rate} follows by similar reasoning to the conclusion of Theorem~\ref{thm:sobolev_testing_rate}.
-
-\section{Proofs of Technical Results}
-\label{sec:technical_lemma_proofs}
-
-Here we give in full detail the proofs of Lemma~\ref{lem:roughness_functional_expectation_sobolev}, \eqref{eqn:spectral_similarity}, and Lemma~\ref{lem:pi_max_pf_1}.
-
-\subsection{Proof of Lemma~\ref{lem:roughness_functional_expectation_sobolev}}
-To simplify exposition, we introduce the iterated difference operator, defined recursively as
-\begin{equation*}
-D_{jk}f(x) = (D_{k}f(x_j) - D_{k}f(x))\frac{K_r(x_j,x)}{r^d},~~ D_jf(x) = (f(x_j) - f(x))\frac{K_r(x_j,x)}{r^d}~~ \textrm{for $j \in [n], k \in [n]^q$}
-\end{equation*}
-We will also use the notation $d_jf(x) := (f(x_j) - f(x))$. We split our analysis into cases based on whether $s$ is even or odd.
-
-\subsubsection{Case 1: $s$ is even.}
-When $s$ is even, letting $q = s/2$ we have the decomposition
-\begin{equation}
-\label{eqn:continuous_to_discrete_sobolev_norm_pf1}
-f^T L^s f =  r^{ds} \cdot \sum_{i = 1}^{n} \sum_{k \in [n]^q} \sum_{\ell \in [n]^q} D_kf(x_i) D_{\ell}f(x_i). 
-\end{equation}
-
-For given index vectors $k,\ell \in [n]^q$ and indices $i,j$, let $I = \abs{k \cup \ell \cup i}$ be the total number of unique indices.
-We separate our analysis into cases based on the magnitude of $I$, specifically whether $I = s + 1$ (the leading terms where all indices are distinct) $I < s + 1$ (the terms where at least one index is repeated) and show that
-\begin{equation}
-\label{eqn:expected_difference_operators_1}
+	\Ebb\bigl[R_{s,n}(f)\bigr] \leq c \cdot \norm{f}_{W_d^{s,2}(\Xset)}
+	\end{equation}
+	for some constant $c$ which may depend on $s$, $p_{\max}$, $K_{\max}$, $d$ and $\Xset$, but not on $f$, $r$ or $n$.
+\end{lemma}
+From Lemma~\ref{lem:roughness_functional_expectation_sobolev}, the result~\eqref{eqn:higher_order_sobolev_testing_rate_pf1} follows by Markov's inequality.
+
+Note that the bound~\eqref{eqn:roughness_functional_expectation_sobolev} involves, on the right hand side, the norm $\norm{f}_{W_d^{s,2}(\Xset)}$ as opposed to the seminorm $[f]_{W_d^{s,2}(\Xset)}$.
+To better understand this, consider the following operator
+\begin{equation*}
+L_rf(x) = \int (f(z) - f(x))K(z,x) \,dP(x)
+\end{equation*} 
+The operator $L_r$ is the expectation of the graph Laplacian, in the sense that $\Ebb(L_nf(x)) = L_rf(x)$, and so it makes sense that the behavior of the associated seminorm $\dotp{L_r^sf}{f}$ is related to the behavior of $\dotp{L_n^sf}{f}$.
+If $p$ is not uniform, for any $s$ the only functions obviously in the null space of $L_r^s$ are constant functions.
+Therefore, the magnitude of each derivative $f^{(\alpha)}$ of $f$ is relevant to the overall expected graph Sobolev seminorm.
+
+To be clear, $\Ebb[\dotp{L_n^sf}{f}] \neq \dotp{L_r^sf}{f}$, and bounding the former turns out to be non-trivial.
+Recall the graph difference operator $D_k$ for $k = (k_1,\ldots,k_n) \in [n]^s$, defined to satisfy the recursive relation
+\begin{equation*}
+D_{k_1}f(x) = (f(x_{k_1}) - f(x))K_r(x_{k_1},x),~~ D_kf(x) = \bigl(D_{k_1}(D_{(k_2,\ldots,k_n)}f\bigr)(x)
+\end{equation*}
+where $K_r(x,z) = \frac{1}{r^d}K(\norm{x - z}/r)$. 
+Then when $s$ is even, letting $q = s/2$ we have
+\begin{equation}
+\label{eqn:roughness_functional_representation_even}
+R_{s,n}(f) = \frac{1}{n}\sum_{i = 1}^{n} \left(\frac{1}{(nr^2)^q}\sum_{k \in [n]^q} D_kf(x_i)\right)^2
+\end{equation}
+and when $s$ is odd, letting $q = (s - 1)/2$ we have
+\begin{equation}
+\label{eqn:roughness_functional_representation_odd}
+R_{s,n}(f) =  \frac{1}{2n^2r^2}\sum_{i,j = 1}^{n}\left(\frac{1}{(nr^2)^q}\sum_{k \in [n]^q}\bigl(D_kf(x_i) - D_kf(x_j)\bigr)\right)^2K_r(x_i,x_j)
+\end{equation} The proof of Lemma~\ref{lem:roughness_functional_expectation_sobolev} therefore relies heavily on Lemma~\ref{lem:expected_difference_operators_sobolev}.
+
+\begin{lemma}
+	\label{lem:expected_difference_operators_sobolev}
+	Let $\Xset$ be a Lipschitz domain, and suppose $f \in W_d^{s,2}(\Xset)$ for some $s \in \mathbb{N}_{+}$.
+Let $q = s/2$ when $s$ is even, and $q = (s - 1)/2$ when $s$ is odd.
+For any indices $k = (k_1,\ldots,k_q)$ and $\ell = (\ell_1,\ldots,\ell_q)$, we have
+	\begin{equation}
+	\label{eqn:expected_difference_operators_sobolev_1}
 	\Ebb(D_kf(x_i) D_\ell f(x_i)) =
 	\begin{cases*}
-O(r^{2s}) \cdot \norm{f}_{W^{s,2}(\Xset)}^2, & ~~\textrm{if $I = s + 1$} \\
-O(r^{2} r^{d(I - (s + 1))}) \cdot [f]_{W^{1,2}(\Xset)}^2, & ~~\textrm{if $I < s + 1$}~ 
+	O(r^{2s}) \cdot \norm{f}_{W_d^{s,2}(\Xset)}^2, & ~~\textrm{if all indices are distinct} \\
+	O(r^{2} r^{d(\abs{k \cup \ell \cup i} - (2q + 1))})\cdot [f]_{W_d^{1,2}(\Xset)}^2, & ~~\textrm{otherwise}~ 
 	\end{cases*}
 	\end{equation}
-We will prove~\eqref{eqn:expected_difference_operators_1} in Section~\ref{subsec:expected_difference_operators_pf}.
-First, we verify that~\eqref{eqn:continuous_to_discrete_sobolev_norm_pf1} and \eqref{eqn:expected_difference_operators_1} are together enough to show Lemma~\ref{lem:roughness_functional_expectation_sobolev} when $s$ is even.
-In the sum on the right hand side of~\eqref{eqn:continuous_to_discrete_sobolev_norm_pf1}, there are $O(n^{I})$ terms with exactly $I$ distinct indices.
-When $I < s + 1$, by~\eqref{eqn:expected_difference_operators_1} the total contribution of such terms to the sum is $O(n^{I}r^{d(I - 1) + 2}) \cdot [f]_{W^{1,2}(\Xset)}^2$.
-Since by assumption $r \geq n^{-1/d}$, this increases with $I$.
-Taking $I = s$ to be the largest integer less than $s + 1$, the contribution of these terms to the sum is therefore $O(n^sr^{d(s - 1) + 2}) \cdot [f]_{W^{1,2}(\Xset)}^2$ which in light of the restriction $r \geq n^{-1/(2(s - 1) + d)}$ is $O(n^{s+1}r^{s(d +2)}) \cdot [f]_{W^{1,2}(\Xset)}^2$.
-On the other hand when $I = s + 1$, by~\eqref{eqn:expected_difference_operators_1} we immediately have that the total contribution of these terms is $O(n^{s + 1}r^{2(s + d)}) \cdot \norm{f}_{W^{s,2}(\Xset)}$.
-Therefore,
-\begin{equation*}
-\Ebb(f^T L^s f) = O(n^{s+1}r^{s(d+2)}) \cdot \norm{f}_{W^{s,2}(\Xset)}.
-\end{equation*}
-
-\subsubsection{Case 2: $s$ is odd.}
-When $s$ is odd, letting $q = (s - 1)/2$ we have
-\begin{equation}
-\label{eqn:roughness_functional_representation_odd}
-f^T L^s f =  r^{ds} \cdot \sum_{i,j = 1}^{n} \sum_{k \in [n]^q} \sum_{\ell \in [n]^q} \bigl(d_jD_kf(x_i)\bigr) \cdot  \bigl(d_jD_{\ell}f(x_i)\bigr) \cdot K_r(x_i,x_j).
-\end{equation}
-For given index vectors $k,\ell \in [n]^q$ and indices $i,j \in [n]$, let $I = \abs{k \cup \ell \cup i \cup j}$ be the total number of unique indices.
-Similar to the case when $s$ is even, we show that
-\begin{equation}
-\label{eqn:expected_difference_operators_2}
+	Additionally, we have
+	\begin{equation}
+	\label{eqn:expected_difference_operators_sobolev_2}
 	\Ebb(d_iD_kf(x_j) d_iD_\ell f(x_j)) =
 	\begin{cases*}
-O(r^{2s}) \cdot \norm{f}_{W_d^{s,2}(\Xset)}^2, & ~~\textrm{if $I = s + 1$} \\
-O(r^{2} \cdot r^{d(I - (s + 1))}) \cdot [f]_{W_d^{1,2}(\Xset)}^2, & ~~\textrm{if $I < s + 1$}~ 
+	O(r^{2s}) \cdot \norm{f}_{W_d^{s,2}(\Xset)}^2, & ~~\textrm{if all indices are distinct} \\
+	O(r^{2} \cdot r^{d(\abs{k \cup \ell \cup i \cup j} - (2q + 2))}) \cdot [f]_{W_d^{1,2}(\Xset)}^2, & ~~\textrm{otherwise}~ 
 	\end{cases*}
 	\end{equation}
-Then Lemma~\ref{lem:roughness_functional_expectation_sobolev} follows from similar reasoning to the case where $s$ was even.
-
-\subsubsection{Proof of~\eqref{eqn:expected_difference_operators_1} and~\eqref{eqn:expected_difference_operators_2}}
-\label{subsec:expected_difference_operators_pf}
+\end{lemma}
+
+The lower bound on $r(n)$ in Lemma~\ref{lem:roughness_functional_expectation_sobolev} is needed to ensure the leading term in~\eqref{eqn:expected_difference_operators_sobolev_1} (or~\eqref{eqn:expected_difference_operators_sobolev_2}), where all indices are distinct, dominates the lower-order terms, where indices are repeated.
+
+\section{Proof of Lemma~\ref{lem:expected_difference_operators_sobolev}}
 
@@
 and additionally $[g]_{W^{\ell,2}(\Rd)} \leq c [f]_{W^{\ell,2}(\Rd)}$ for each $\ell = 0,\ldots,s$.
-Then if \eqref{eqn:expected_difference_operators_1} and \eqref{eqn:expected_difference_operators_2} hold with respect to $g$, they hold (up to constants) with respect to $f$ as well.
-
-To construct such a $g$, we first take an extension of $f$ be defined on all of $\Rd$, and then mollify.
+Then if \eqref{eqn:expected_difference_operators_sobolev_1} and \eqref{eqn:expected_difference_operators_sobolev_2} hold with respect to $g$, they hold (up to constants) with respect to $f$ as well.
+
+To construct such a $g$, we first take an extension of $f$ to be defined over $\Rd$, and then mollify.
 Since $\Xset$ is a Lipschitz domain, there exists an extension \textcolor{red}{citation} $\wt{f}$ of $f$ compactly supported on $\Rd$ such that $\wt{f} = f$ a.e.
@@
 
-Now, since $\wt{f} \in W^{s,2}(\Rd)$, there exists a sequence $(g_m) \subset C^{\infty}(\Rd)$ such that $\norm{g_m - \wt{f}}_{W^{s,2}(\Rd)} \to 0$. On the one hand, by the Cauchy-Schwarz inequality
-\begin{equation*}
-\abs{\Ebb\bigl[D_k\wt{f}(x_i) D_{\ell} \wt{f}(x_i)\bigr] - \Ebb\bigl[D_kg_m(x_i) D_\ell g_m(x_i)\bigr]} \leq \frac{c}{r^{sd}} \cdot \norm{f - g_m}_{\Leb^2(\Rd)}
+Now, since $\wt{f} \in W^{s,2}(\Rd)$, there exists a sequence $(g_m) \in C^{\infty}(\Rd)$ such that $\norm{g_m - \wt{f}}_{W^{s,2}(\Rd)} \to 0$. On the one hand, by the Cauchy-Schwarz inequality
+\begin{equation*}
+\abs{\Ebb[D_k\wt{f}(x_i) D_{\ell} \wt{f}(x_i)] - \Ebb[D_kg_m(x_i) D_\ell g_m(x_i)]} \leq \frac{c}{r^{sd}} \cdot \norm{f - g_m}_{\Leb^2}(\Rd) 
 \end{equation*}
 and taking $m$ to be sufficiently large, we can make the right hand side less than $\delta$.
-On the other hand, since $\norm{g_m - \wt{f}}_{W^{s,2}(\Rd)} \to 0$, there exists $m$ sufficiently large such that $[g_m]_{W^{s,2}(\Rd)}$ is at most say $2 [\wt{f}]_{W^{s,2}(\Rd)}$.
-We therefore take $g = g_m$ for $m$ large enough to satisfy both conditions.
-
-Our task is now to prove that \eqref{eqn:expected_difference_operators_1} and \eqref{eqn:expected_difference_operators_2} hold with respect to $g$.
+On the other hand, since $\norm{g_m - \wt{f}}_{W^{s,2}(\Rd)} \to 0$, there exists $m$ sufficiently large such that $[g_m]_{W^{s,2}(\Rd)} \leq 2 [\wt{f}]_{W^{s,2}(\Rd)}$.
+We take $m$ sufficiently large such that $g = g_m$ satisfies both conditions.
+
+Our task is now to prove that \eqref{eqn:expected_difference_operators_sobolev_1} and \eqref{eqn:expected_difference_operators_sobolev_2} hold with respect to $g$.
 We first prove the desired bounds in the case when some indices are repeated, and then the desired bounds in the case when all indices are distinct.
 
-\paragraph{Repeated indices.}
-
-Since the proofs of~\eqref{eqn:expected_difference_operators_1} and~\eqref{eqn:expected_difference_operators_2} are essentially the same for the case where some index is repeated, we will assume without loss of generality that $s$ is even.
+\subsubsection{Repeated indices.}
+
+Since the proofs of~\eqref{eqn:expected_difference_operators_sobolev_1} and~\eqref{eqn:expected_difference_operators_sobolev_2} are essentially the same for the case where some index is repeated, we will assume without loss of generality that $s$ is even.
 Let $k,\ell \in [n]^q$ be index vectors for $q = s/2$.
@@
 
-\textit{Case 1: Two distinct indices.}
+\paragraph{Case 1: Two distinct indices.}
 Let $k_q = \ell_q = i$, and $h = j$. Using the law of iterated expectation, we obtain
@@
 where the last equality follows from Lemma~\ref{lem:expected_first_order_seminorm}.
-
-\textit{Case 2: Three distinct indices.}
+\paragraph{Case 2: Three distinct indices.}
 Let $k_q = \ell_q = i$, for some $i \neq j \neq h$. Using the law of iterated expectation, we obtain
@@
 
-\textit{Case 3: Four distinct indices.}
+\paragraph{Case 3: Four distinct indices.}
 Using the law of iterated expectation, we find that
@@
 \end{align*}
-where the last equality follows from Lemma~\ref{lem:expected_first_order_seminorm_3}.
-
-\paragraph{All indices distinct.}
+where the last inequality follows from Lemma~\ref{lem:expected_first_order_seminorm_3}.
+
+\subsubsection{All indices distinct.}
 
@@
 
-\textit{Case 1: $s$ is even.}
-
-By Lemma~\ref{lem:leading_term_sobolev} there exists some $f_s \in \Leb^2(\Rd)$ which satisfies $\norm{f_s}_{\Leb^2(\Rd)} \leq c \norm{g}_{W^{s,2}(\Rd)}$ such that
-\begin{equation*}
-\Ebb\Bigl[\bigl(D_kg(x)\bigr)^2\Bigr] = r^{2s} \cdot \norm{f_s}_{\Leb^2(\Rd)}.
-\end{equation*} Therefore, by the law of iterated expectation along with this Lemma, 
-\begin{align*}
-\Ebb\bigl[D_kg(x_i)D_kg(x_j)\bigr] = \Ebb\Bigl[\bigl(\Ebb[D_kg(x_i)|x_i]\bigr)^2\Bigr] = r^{2s} \cdot \Ebb\Bigl[\bigl(f_s(x_i)\bigr)^2\Bigr] = r^{2s} \cdot \norm{g}_{W^{s,2}(\Rd)},
+\paragraph{Case 1: $s$ even.}
+
+By Lemma~\ref{lem:leading_term_sobolev} there exists some $f_s \in \Leb^2(\Rd)$ which satisfies $\norm{f_s}_{\Leb^2(\Rd)} \leq c \norm{g}_{W^{s,2}(\Rd)}$.
+Therefore, by the law of iterated expectation along with this Lemma,
+\begin{align*}
+\Ebb\bigl[D_kg(x_i)D_kg(x_j)\bigr] = \Ebb\Bigl[\bigl(\Ebb[D_kg(x_i)|x_i]\bigr)^2\Bigr] = O(r^{2s})\cdot \Ebb\Bigl[\bigl(f_s(x_i)\bigr)^2\Bigr] = O(r^{2s}) \cdot \norm{g}_{W^{s,2}(\Rd)},
 \end{align*}
@@
 
-\textit{Case 2: $s$ is odd.}
+\paragraph{Case 2: $s$ odd.}
 
@@
 \Ebb[d_iD_kg_m(x_j)d_iD_{\ell}g_m(x_j)K_r(x_i,x_j)] & = \Ebb\biggl[\Bigl(d_i\bigl(\Ebb(D_kf)\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \\
-& = \Ebb\biggl[\Bigl(d_i\bigl(r^{s - 1}\cdot f_{s - 1} + r^sf_s(x_j)\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr].
-\end{align*}
-where the latter equality follows from Lemma~\ref{lem:leading_term_sobolev}. Here $f_{s -1}$ and $f_s$ satisfy the conclusions of that Lemma, namely that $f_{s - 1}$ and $f_s \in C^{\infty}(\Rd)$, and
+& = \Ebb\biggl[\Bigl(d_i\bigl(I_{s - 1}\cdot f_{s - 1} + O(r^s)f_s(x_j)\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr].
+\end{align*}
+where the latter equality follows from Lemma~\ref{lem:leading_term_sobolev}.
+Here $I_{s - 1}, f_{s -1}$ and $f_s$ satisfy the conclusions of that Lemma, namely that $\abs{I_{s - 1}} \leq r^{s - 1}$, $f_{s - 1}$ and $f_s \in C^{\infty}(\Rd)$, and
 \begin{equation*}
@@
 \end{equation*}
-By the linearity of the difference operator $d_i$, we have
-\begin{align*}
-\Ebb\biggl[\Bigl(d_i\bigl(r^{s - 1}\cdot f_{s - 1} + r^s f_s(x_j)\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] & = \Ebb\biggl[\Bigl(r^{s - 1} \cdot d_if_{s - 1}(x_j) + r^s d_if_s(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \\
-& \leq 2r^{2(s - 1)} \Ebb\Bigl[\bigl(d_if_{s-1}(x_j)\bigr)^2K_r(x_i,x_j)\Bigr] + 2r^{2s}\Ebb\Bigl[\bigl(d_if_s(x_j)\bigr)^2\Bigr] \\
-& \leq 2r^{2(s - 1)} \Ebb\Bigl[\bigl(d_if_{s-1}(x_j)\bigr)^2K_r(x_i,x_j)\Bigr] + 4r^{2s} \norm{g}_{W^{s,2}(\Rd)}^2 \\
-& \leq r^{2s} \cdot c\norm{g}_{W^{s,2}(\Rd)}^2
+By the linearity and boundedness of the difference operator $d_i$, we have
+\begin{align*}
+\Ebb\biggl[\Bigl(d_i\bigl(I_{s - 1}\cdot f_{s - 1} + O(r^s)f_s(x_j)\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] & = \Ebb\biggl[\Bigl(I_{s - 1} d_if_{s - 1}(x_j) + O(r^s)f_s(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \\
+& \leq 2I_{s - 1}^2 \Ebb\Bigl[\bigl(d_if_{s-1}(x_j)\bigr)^2K_r(x_i,x_j)\Bigr] + O(r^{2s})\Ebb\Bigl[f_s(x_j)^2\Bigr] \\
+& = O(r^{2(s - 1)})\Ebb\Bigl[\bigl(d_if_{s-1}(x_j)\bigr)^2K_r(x_i,x_j)\Bigr] + O(r^{2s}) \norm{g_m}_{W^{s,2}(\Rd)}^2 \\
+& \leq O(r^{2s}) \cdot \norm{g_m}_{W^{s,2}(\Rd)}^2
 \end{align*}
 where the last inequality follows from Lemma~\ref{lem:expected_first_order_seminorm}.
-This concludes the proof of~\eqref{eqn:expected_difference_operators_1} and~\eqref{eqn:expected_difference_operators_2}.
-
-\subsection{Proof of~\eqref{eqn:spectral_similarity}}
-
-Assuming~\eqref{eqn:slepcev_transport_distance} holds and $r$ is chosen according to~\eqref{eqn:pi_max}, we have already shown that the upper bound in~\eqref{eqn:spectral_similarity} holds for $\delta = 0$ (see the proof of ~\eqref{eqn:eigenvalue_tail_bound}).
-Now we need only to show the lower bound, which we will do following these steps:
-\begin{enumerate}[(i)]
-	\item First, we consider the case of two fixed graphs $G$ and $\wt{G}$, and establish a Poincare inequality.
-In other words, we show that if edges in $G$ can be mapped to short, non-overlapping paths in $\wt{G}$, then the Laplacian matrix $L$ cannot be much larger $\wt{L}$, in the usual sense of partial ordering of matrices.
-	\item Then, we treat the case of $G = G_{n,r}$ and $\wt{G} = \overline{G}$.
-We assume that~\eqref{eqn:slepcev_transport_distance} holds, and exhibit such a mapping between edges in $G_{n,r}$ and paths in $\overline{G}$.
-The resulting Poincare inequality will imply~\eqref{eqn:spectral_similarity}.
-\end{enumerate}
-In step (ii) we assume that~\eqref{eqn:slepcev_transport_distance} holds.
-Since \eqref{eqn:slepcev_transport_distance} holds with probability at least $1 - c/n$, this will imply that~\eqref{eqn:spectral_similarity} holds with at least the same probability.
-
-\subsubsection{Step 1: Poincare inequality}
-
-Let $G = (V,E)$ and $\wt{G} = (V,\wt{E})$ be two graphs over a common vertex set $V$. The set $\mathcal{P}_{\wt{G}}$ consists of all paths $P$ over $\wt{G}$, meaning all tuples
-\begin{equation*}
-P = (\wt{e}_1,\wt{e}_2,\ldots,\wt{e}_m),~~m \in \mathbb{N}~, \wt{e}_j \in \wt{E}~ \forall~ j \in [m], (\wt{e}_j)_2 = (\wt{e}_{j + 1})_1 \forall~j \in [m - 1].
-\end{equation*}
-Let $\gamma:E \to \mathcal{P}_{\wt{G}}$ be a mapping from edges in $G$ to paths in $\wt{G}$. The maximum path length $M(\gamma)$ is defined
-\begin{equation*}
-M(\gamma) = \max_{e \in E} \abs{\gamma(e)}
-\end{equation*} 
-and the bottleneck $b(\gamma)$ is defined
-\begin{equation*}
-b(\gamma) = \max_{\wt{e} \in \wt{E}} \abs{\{e \in E:  \wt{e} \in \gamma(e)\}}.
-\end{equation*}
-As mentioned previously, the Laplacian matrix $L$ cannot be much larger than $\wt{L}$ if edges in $G$ can be mapped to short, non-overlapping paths in $\wt{G}$.
-Formally, we will show the following: for any $\gamma: E \to \mathcal{P}_{\wt{E}}$,
-\begin{equation}
-\label{eqn:poincare_inequality}
-x^T L x \leq M(\gamma)\cdot b(\gamma) \cdot (x^T \wt{L} x)~\forall{x \in \Reals^n}
-\end{equation}
-
-\paragraph{Proof of~\eqref{eqn:poincare_inequality}:}
-
-Letting $c \in \Reals$, we will use the notation $G \preceq c \cdot  \wt{G}$ as shorthand for
-\begin{equation*}
-x^T L x \leq c \cdot  x^T \wt{L} x,~~\textrm{for all $x \in \Reals^n$}.
-\end{equation*}
-
-Let $G_e = (V, \set{e})$ and $P_e = (V, \set{\widetilde{e}: \widetilde{e} \in \gamma(e)})$ be the graphs associated with $e$ and $\gamma(e)$, respectively. By Lemma \ref{lem: path_poincare}, we have
-\begin{equation*}
-G_{e} \preceq \abs{P_e} P_e
-\end{equation*}
-Summing over all $e \in E_G$, we obtain
-\begin{align*}
-G & \preceq \sum_{e \in E_G} \abs{P_e} P_e \\
-& \preceq \max_{e \in E_G} \abs{\gamma(e)} \sum_{e \in E_G} P_e \\
-& \preceq \max_{e \in E_G} \abs{\gamma(e)} b_{\gamma}\cdot \widetilde{G}
-\end{align*}
-
-\subsubsection{Step 2: Mapping $E_{G_{n,r}}$ to $\mathcal{P}_{\overline{G}}$}
-We assume there exists a mapping $\pi:\overline{X} \to X$ which satisfies \eqref{eqn:slepcev_transport_distance}.
-We will apply the Poincare inequality to $\pi^{-1}(G_{n,r})$, the isomorphism of $G_{n,r}$ induced by the mapping of vertices $\pi^{-1}: X \to \overline{X}$.
-We will map a given $(\overline{x}_{k},\overline{x}_{\ell}) \in \pi^{-1}(E)$ to a shortest path $P \in \mathcal{P}_{\overline{G}}$ (measured by Manhattan distance) between $\overline{x}_k$ and $\overline{x}_{\ell}$.
-Recalling the notation $t = n^{1/d}$ and letting
-\begin{equation*}
-s_i = \frac{1}{t}(\textrm{sign}(k_i - \ell_i),0,\ldots,0),~~\textrm{for $i = 1,\ldots,d$,}
-\end{equation*}
-our mapping $\gamma: \pi^{-1}(E) \to \mathcal{P}_{\overline{G}}$ is given by
-\begin{align*}
-\gamma((\overline{x}_{k},\overline{x}_{\ell})) = \bigl(& (\overline{x}_k, \overline{x}_{k} + s_1), (\overline{x}_{k} + s_1,\overline{x}_{k} + 2s_1),\ldots,(\overline{x}_{k} + (\abs{k_1 - \ell_1} - 1)s_1, \overline{x}_{k} + \frac{1}{t}(k_1 - \ell_1)) \\
-& (\overline{x}_{k} + \frac{1}{t}(k_1 - \ell_1), \overline{x}_{k} + \frac{1}{t}(k_1 - \ell_1) + s_2),\ldots,(\overline{x}_{k} + \frac{1}{t}(k_1 - \ell_1) + (\abs{k_2 - \ell_2} - 1)s_2, \overline{x}_{k}  + \frac{1}{t}((k_1 - \ell_1) + (k_2 - \ell_2)) \\
-& \vdots \\
-& (\overline{x}_{k} + \frac{1}{t}\sum_{i = 1}^{d - 1}(k_i - \ell_i) + s_d,\overline{x}_{k} + \frac{1}{t}\sum_{i = 1}^{d - 1}(k_i - \ell_i) + 2s_d), \ldots, (\overline{x}_k + \frac{1}{t}\sum_{i = 1}^{d - 1}(k_i - \ell_i) + (\abs{k_d - \ell_d} - 1)s_d, \overline{x}_{\ell})\bigr)
-\end{align*}
-As a result of~\eqref{eqn:transport_distance} we obtain the following bounds on $M(\gamma)$ and $b(\gamma)$:
+
+\subsection{Additional Lemmas}
+
+We use multiindex notation to represent higher order partial derivatives and polynomials. For $\alpha \in [\mathbb{N}]^d$, and $x,z \in \Rd$ we write
+\begin{equation*}
+\abs{\alpha} = \alpha_1 + \cdots + \alpha_d,~~ f^{(\alpha)} = \frac{\partial^{\abs{\alpha}}f}{\partial x_{\alpha_1}\ldots \partial x_{\alpha_q}},~~ (x - z)^{\alpha} := (x_{\alpha_1} - z_{\alpha_1})\cdots(x_{\alpha_d} - z_{\alpha_d})
+\end{equation*}
+
+\begin{lemma}
+	\label{lem:leading_term_sobolev}
+	Let $k \in [n]^q$ for some $q \geq 1$. Suppose that $g \in C^{\infty}(\Rd)$ and $p \in C^{s - 1}(\Rd;p_{\max})$, and that $K_r$ is a second order kernel. Then there exist
+	\begin{itemize}
+		\item functions $f_{\ell}, \ell = 2q,\ldots,s$ satisfying $f_{\ell} \in W_d^{s - \ell,2}(\Rd) \cap C_d^{\infty}(\Rd)$ and
+		\begin{equation*}
+		\norm{f_{\ell}}_{W_d^{s - \ell,2}(\Rd)} \leq c \norm{f}_{W_d^{s,2}(\Rd)}
+		\end{equation*}
+		for some constant $c$ which depends only on $d$, $\Xset$ and $p_{\max}$, and 
+		\item constants $I_{\ell},\ell = 2q,\ldots,s$ which depend only on $K(\cdot)$ satisfying $\abs{I_{\ell}} \leq r^{\ell}$,
+	\end{itemize}
+	such that
+	\begin{equation}
+	\label{eqn:leading_term_sobolev}
+	\Ebb(D_kf(x)) =
+	\begin{cases*}
+	\sum_{\ell = 2q}^{s - 1} I_{\ell} \cdot f_{\ell}(x) +  O(r^{s}) \cdot f_{s}(x),~~& \textrm{if $2q < s$} \\
+	O(r^{s}) \cdot f_{s}(x),~~& \textrm{if $2q \geq s$}
+	\end{cases*}
+	\end{equation}
+	The $O(\cdot)$ term may depend on $s, K_{\max}$ and $p_{\max}$, but does not depend on $f$.
+\end{lemma}
+\begin{proof}	
+	We proceed by induction on $q$. 
+	\paragraph{Base case.}
+	We begin with the base case of $q = 1$. Since $f$ and $p$ are smooth, they both admit Taylor expansions of the following form for all $x,z \in \Rd$:
+	\begin{align*}
+	f(z) & = \sum_{\abs{\alpha} < s} \frac{f^{(\alpha)}(x)}{\alpha!} (x - z)^{\alpha} + \frac{\abs{\alpha}}{\alpha!}\sum_{\abs{\alpha} = s} (x - z)^{\alpha} \int_{0}^{1}(1 - t)^{s - 1} f^{(\alpha)}(x + t(z - x)) \,dt\\
+	p(z) & = \sum_{\abs{\beta} < s  - 1} \frac{p^{(\beta)}(x)}{\beta!} (x - z)^{\beta} + O((x - z)^{s - 1})
+	\end{align*}
+	where $f^{(\alpha)} \in W^{s - \abs{\alpha},2}(\Rd) \cap C^{\infty}(\Rd)$ additionally satisfies
+	\begin{equation*}
+	\norm{f^{(\alpha)}}_{W^{s - \abs{\alpha},2}(\Rd)} \leq \norm{f}_{W^{s,2}(\Rd)}
+	\end{equation*}
+
+	Replacing $f$ by its Taylor expansion inside the expected first order difference operator $\Ebb(D_kf(x))$ and letting $E_{\alpha,P} := \Ebb\left[(x - x_k)^{\alpha}K_r(x_k,x)\right]$, we have
 	\begin{align}
-M(\gamma) & \leq c n^{1/d} \left(\left(\frac{\log n}{n}\right)^{1/d} + r\right) \label{eqn:maximum_path_length}\\
-b(\gamma) & \leq c \left(n^{1/d} \left(\left(\frac{\log n}{n}\right)^{1/d} + r\right)\right)^{2d}, \label{eqn:bottleneck}
+	\Ebb(D_kf(x)) & = \sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)}{\alpha!} E_{\alpha,P} + \frac{\abs{\alpha}}{\alpha!}\sum_{\abs{\alpha} = s} \int_{0}^{1}(1 - t)^{s - 1} \Ebb\bigl[f^{(\alpha)}(x + t(x_k - x)) (x_k - x)^{\alpha} K_r(x_k,x)\bigr] \,dt \nonumber \\
+	& =  \sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)}{\alpha!} E_{\alpha,P} + O(r^s)\sum_{\abs{\alpha} = s} \int_{0}^{1}\Ebb\bigl[f^{(\alpha)}(x + t(x_k - x))K_r(x_k,x)\bigr] \,dt \nonumber \\
+	& =  \sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)}{\alpha!} E_{\alpha,P} + O(r^s)\sum_{\abs{\alpha} = s} \int_{0}^{1} \int_{B(x,r)}\frac{f^{(\alpha)}(x + t(z - x))}{r^d} \,dz \,dt \nonumber \\
+	& = \sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)}{\alpha!} E_{\alpha,P} + O(r^s)\int_{0}^{1} \int_{B(0,1)} \sum_{\abs{\alpha} = s} f^{(\alpha)}(x + try) \,dy \,dt \label{eqn:leading_term_sobolev_pf1}
 	\end{align}
-which we now prove.
-
-\paragraph{Proof of~\eqref{eqn:maximum_path_length}:}
-
-Suppose $(\overline{x}_k, \overline{x}_{\ell}) \in \pi^{-1}(E)$. Then
-\begin{align*}
-\gamma(\overline{x}_k, \overline{x}_{\ell})) = \norm{\overline{x}_k - \overline{x}_{\ell}}_1 & \leq \sqrt{d} \norm{\overline{x}_k - \overline{x}_{\ell}}_2 \\
-& \leq \sqrt{d}\bigl(\norm{\overline{x}_k - \pi(\ol{x}_{k})}_2 + \norm{\pi(\ol{x}_k) - \pi(\ol{x}_{\ell})}_2 + \norm{\overline{x}_\ell - \pi(\ol{x}_{\ell})}_2\bigr) \\
-& \leq \sqrt{d}\left(\norm{\pi(\ol{x}_k) - \pi(\ol{x}_{\ell})}_2 + 2c\left(\frac{\log n}{n}\right)^{1/d}\right) \tag{by \eqref{eqn:transport_distance}} \\
-& \leq \sqrt{d}\left(r + 2c\left(\frac{\log n}{n}\right)^{1/d}\right) \tag{since $(\pi(\ol{x}_k),\pi(\ol{x}_{\ell})) \in E$}
-\end{align*}
-
-\paragraph{Proof of~\eqref{eqn:bottleneck}:}
-
-Suppose $(\overline{x}_k, \overline{x}_{\ell}) \in \gamma(e)$ for some $e = (\overline{x}_i, \overline{x}_j) \in \pi^{-1}(E)$.  Then, by the construction of $\gamma$,
-\begin{equation}
-\label{eqn:bottleneck_pf_1}
-\norm{\overline{x}_k - \overline{x}_i}_2 \wedge \norm{\overline{x}_\ell - \overline{x}_j}_2  \leq \norm{\overline{x}_j - \overline{x}_i}_2
-\end{equation}
-As shown in the preceding paragraph, assuming~\eqref{eqn:transport_distance} we have
-\begin{equation*}
-\norm{\overline{x}_j - \overline{x}_i}_2 \leq r + 2c\left(\frac{\log n}{n}\right)^{1/d}
-\end{equation*}
-and therefore by~\eqref{eqn:bottleneck_pf_1},
-\begin{align*}
-\abs{\set{e \in E: (\overline{x}_k, \overline{x}_{\ell}) \in \gamma(e)}} & \leq \abs{\set{\overline{x}_i: \norm{\overline{x}_k - \overline{x}_i}_2 \leq r + 2c\left(\frac{\log n}{n}\right)^{1/d}}} \cdot \abs{\set{\overline{x}_i: \norm{\overline{x}_k - \overline{x}_i}_2 \leq r + 2c\left(\frac{\log n}{n}\right)^{1/d}}} \\
-& \leq \left(n^{1/d}\left( r + 2c\left(\frac{\log n}{n}\right)^{1/d}\right)\right)^2.
-\end{align*}
-
-Putting the pieces together, we have that when $d = 1$ and $r = (\log n)^a (\log n/n)$, then
-\begin{equation*}
-G_{n,r} \preceq c \cdot (\log n)^{(1/d + a) + 2 + 2da}\cdot \overline{G}
-\end{equation*}
-with probability at least $1 - o(1)$.
-
-\subsection{Proof of Lemma~\ref{lem:pi_max_pf_1}}
-
-Write $\overline{L} = \overline{V}\overline{S} \overline{V}^T$ for the spectral decomposition of $\overline{L}$. The proof of Lemma~\ref{lem:pi_max_pf_1} will proceed according to the following steps.
-\begin{enumerate}
-	\item We show that the incoherence $\max_{i} v_{k,i}^2$ can be upper bounded
-	\begin{equation}
-	\label{eqn:incoherence_proof_1}
-	\max_{i} v_{k,i}^2 \leq \frac{1}{n} \left(\sum_{j = 1}^{n} \abs{\dotp{v_k}{\overline{v}_j}}\right)^2
-	\end{equation}
-	replacing the norm $\norm{v_k^2}_{\infty}$ by the inner products $\dotp{v_k}{\overline{v}_j}$.
-	\item Let $I(k,r) = \set{j \geq 0: \abs{j - k} \leq r}$. Using Davis-Kahan, we establish that the following upper bounds
-	\begin{equation}
-	\label{eqn:incoherence_proof_2}
-	\sum_{j \not\in I(k,r)}^{n} (\dotp{v_k}{\overline{v}_j})^2 \leq 400 \frac{n^4 \delta^2 s_{\max}^2}{R^4}
-	\end{equation}
-	hold for each $k \leq n/8$ and any $r \geq 8k\sqrt{\delta}$. 
-	\item We carefully upper bound the $L_1$ norm present in \eqref{eqn:incoherence_proof_1} given the various bounds on $L_2$ norm we've established in \eqref{eqn:incoherence_proof_2}.
-\end{enumerate}
-
-\paragraph{Step 1: Proof of~\eqref{eqn:incoherence_proof_1}.}
-
-We can re-express $v_k$ in the basis $(\overline{v}_j)$ as
-\begin{equation*}
-v_k = \sum_{j = 1}^{n} \dotp{v_k}{\overline{v}_j} \overline{v}_j,
-\end{equation*}
-whereupon the bound~\eqref{eqn:incoherence_proof_1} follows from the incoherence property of the chain
-\begin{equation*}
-\max_{i,k = 1,\ldots,n} \overline{v}_{k,i}^2 \leq \frac{1}{n}.
-\end{equation*}
-\paragraph{Step 2: Proof of~\eqref{eqn:incoherence_proof_2}.}
-
-Let $\widetilde{L} = L + H$. By Davis Kahan, we have that
-\begin{equation}
-\label{eqn:davis_kahan}
-\sum_{j \not\in I(k,r)} (\dotp{v_k}{\overline{v}_j})^2 \leq \left(\frac{\norm{H}_{\textrm{op}}}{\min{\abs{\overline{s}_j - s_k}:j \in I(k,r)}}\right)^2
-\end{equation}
-To upper bound the numerator, we use~\eqref{eqn:spectral_similarity} to get
-\begin{equation*}
-\norm{H}_{op} \leq \delta s_{n}.
-\end{equation*}
-
-To lower bound the denominator, we note
-\begin{align}
-\abs{\overline{s}_j - s_k} & \geq \abs{\overline{s}_j - \overline{s}_k} - \abs{\overline{s}_k - s_k} \nonumber \\
-& \geq \abs{\overline{s}_j - \overline{s}_k} - \delta \overline{s}_k. \label{eqn:incoherence_proof_8}
-\end{align}
-The eigenvalues of the chain graph are well known to be
-\begin{equation*}
-\overline{s}_j = 2\left(1 - \cos\left(\frac{j\pi}{n}\right)\right) = 4\sin^2\left(\frac{k\pi}{n}\right) ~~\textrm{for $k = 0,\ldots,n - 1$.}
-\end{equation*}
-By Taylor expansion, for $k \leq j \leq n/(2\pi)$ we have
-\begin{equation}
-\label{eqn:incoherence_proof_5}
-\overline{s}_j - \overline{s}_k = 2\left(\cos\left(\frac{k\pi}{n}\right) - \cos\left(\frac{j\pi}{n}\right)\right) \geq \frac{(k - j)^2\pi^2}{n^2}.
-\end{equation}
-and when $j \geq n/(2\pi)$ we have $\overline{s}_j - \overline{s}_k \geq 2(\cos\left(\frac{\pi}{8}\right) - \cos\left(\frac{1}{2}\right) > .09$.
-Additionally since $\sin(x) \leq x$ for all $x \geq 0$, we have
-\begin{equation}
-\label{eqn:incoherence_proof_6}
-\overline{s}_k \leq 4\frac{k^2\pi^2}{n^2}.
-\end{equation}
-Combining~\eqref{eqn:incoherence_proof_5},\eqref{eqn:incoherence_proof_6}, and the lower bound $\abs{k - j} \geq R \geq 8k\sqrt{\delta}$, we have that
-\begin{equation}
-\label{eqn:incoherence_proof_7}
-\abs{\overline{s}_j - \overline{s}_k} - \delta \overline{s}_k \geq .04\frac{(k - j)^2\pi^2}{n^2} - 4 \delta\frac{k^2\pi^2}{n^2} \geq .02\frac{(k - j)^2\pi^2}{n^2}.
-\end{equation}
-The result follows from \eqref{eqn:incoherence_proof_7}, \eqref{eqn:incoherence_proof_8} and \eqref{eqn:davis_kahan}.
-
-\paragraph{Step 3: $L_1$ norm to $L_2$ norm.}
-
-Let $u \in \Reals^n$, and suppose we know that the $L_2$ norm of $u$ is bounded,
-\begin{equation}
-\label{eqn:incoherence_proof_3}
-\norm{u}_2^2 \leq 1.
-\end{equation}
-Under no other conditions on $u$, the upper bound $\norm{u}_1 \leq \sqrt{n}$ is the best that can be hoped for (achieved when $u = (n^{-1/2},\ldots,n^{-1/2})$).
-However, suppose we also know that for some $B_2^2 \geq B_3^2 \geq \ldots \geq B_n^2$, we have that there exists some $R \in [n]$ such that
-\begin{equation}
-\label{eqn:incoherence_proof_4}
-\sum_{j = r}^{n} v_j^2 \leq B_r^2 ~~\textrm{for each $r = R,\ldots,n$.}
-\end{equation}
-Clearly, if $B_n^2 \leq \frac{1}{n}$ then $u$ cannot be equal to $(n^{-1/2},\ldots,n^{-1/2})$.
-We might hope for more general improvements on the bound $\norm{u}_1 \leq \sqrt{n}$ if $B_R^2$ is quite small once $R$ gets sufficiently large.
-The following Lemma gives such a result.
-\begin{lemma}
-	\label{lem:pi_max_pf_1_util_1}
-	Suppose $u \in \Reals^n$ satisfies \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4} for some $1 = B_1^2 = B_2^2 \geq \ldots \geq B_n^2 \geq B_{n+1}^2 = 0$.
-Assume additionally that there exists an $R \in [n]$ such that
-	\begin{equation}
-	\label{eqn:incoherence_util_1}
-	B_r^2 - B_{r + 1}^2 \leq \frac{1}{r}~\textrm{for each $r = R, R+1,\ldots, n - 1$.}
-	\end{equation}
-	Then,
-	\begin{equation}
-	\label{eqn:incoherence_util_2}
-	\norm{u}_1 \leq \sqrt{R}\sqrt{1 - B_R^2} + \sum_{j = R}^{n} \sqrt{B_j^2 - B_{j+1}^2}.
-	\end{equation}
-\end{lemma}
-\begin{proof}
-	Set
-	\begin{equation*}
-	(u_\star)_j^2 = 
-	\begin{cases*}
-	\frac{1}{R}(1 - B_R^2), ~ j \leq R \\
-	B_j^2 - B_{j + 1}^2, j > R
-	\end{cases*}
-	\end{equation*}
-	so that $\norm{u_\star}_1$ is equal to the right hand side of~\eqref{eqn:incoherence_util_2}.
-We now prove by contradiction that $u_\star$ maximizes $\norm{\cdot}_1$ under the constraints \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4}.
-Suppose this were not true, and that the maximum is achieved for some $u \neq u_{\star}$.
-	\begin{itemize}
-		\item since $\norm{u}_1 > \norm{u_{\star}}$ then there exists some index $j$ such that
-		\begin{equation}
-		\label{eqn:incoherence_util_3}
-		u_j > (u_{\star})_j
-		\end{equation} 
-		\item by~\eqref{eqn:incoherence_proof_3} $\norm{u}_2 \leq 1$. Since $\norm{u_{\star}}_2 = 1$, by \eqref{eqn:incoherence_util_3} there must exist some index $k$ such that
-		\begin{equation}
-		\label{eqn:incoherence_util_4}
-		(u_{\star})_k > u_k.
-		\end{equation} 
-		Choose $k$ to be the largest of all such indices.
-		\item suppose $k < j$ and $j > R$.
-Since $k$ was chosen to be the largest such index which satisfied~\eqref{eqn:incoherence_util_4}, clearly $v_i \geq (u_{\star})_i$ for all $i > j$, and by \eqref{eqn:incoherence_util_3} $u_j > (u_{\star})_j$.
-But then
-		\begin{equation*}
-		B_j^2 = \sum_{i = j}^{N} (u_{\star})_i^2 < \sum_{i = j}^{N} u_i^2 
-		\end{equation*}
-		and so $v$ does not satisfy~\eqref{eqn:incoherence_proof_4}. Therefore either $k > j$ or $j \leq R$.
-		\item In either case, we have
-		\begin{equation*}
-		u_j > (u_{\star})_j \geq (u_{\star})_k > u_k.
-		\end{equation*}
-		Let $j \leq l < k$ be the largest index for which $v_j > (a_{\star})_j$.
-
-		We now construct a vector $\wt{u}$ which satisfies the constraints \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4} such that $\norm{\wt{u}}_1 > \norm{u}_1$.
-Once we have shown this, we will have established a contradiction, and the proof of Lemma~\ref{lem:pi_max_pf_1_util_1} will be complete.
-Let $\wt{u} = (\wt{u}_i)$ be given as follows:
-		\begin{equation*}
-		\wt{u}_i = 
-		\begin{cases*}
-		u_i, ~~\textrm{if $i \neq k,l$}, \\
-		u_{\star,k}, ~~\textrm{if $i = k$}, \\
-		\sqrt{u_l^2 - (u_{\star,k}^2 - a_k^2)}, ~~\textrm{if $i = l$.}
-		\end{cases*}
-		\end{equation*}
-		Clearly $\norm{\wt{u}}_2 = \norm{u}_2 = 1$ and so $\wt{u}$ satisfies~\eqref{eqn:incoherence_proof_3}.
-To see that $\wt{u}$ satisfies~\eqref{eqn:incoherence_proof_4} we separate the analysis into cases.
-When $r \geq l + 1$, we have that
-		\begin{equation*}
-		\sum_{i = r}^{n} \wt{u}_i^2 = \sum_{i = r}^{n} u_i^2 \leq B_r^2.
-		\end{equation*}
-		When $l < r \leq k$, we have that
-		\begin{equation*}
-		\sum_{i = r}^{n} \wt{u}_i^2 \leq \sum_{i = r}^{k} u_{\star,i}^2 + \sum_{i = k +1}^{n} u_{i}^2 \leq B_r^2 - B_{k + 1}^2 + B_{k + 1}^2 = B_r^2
-		\end{equation*}
-		Finally, when $R \leq r \leq l$, we have
-		\begin{equation*}
-		\sum_{i = r}^{n} \wt{u}_i^2 \leq u_l^2 - (u_{\star,k}^2 - u_k^2) + (u_{\star,k}^2) + \sum_{i \neq k,l} u_i^2 = \sum_{i = r}^{n} u_i^2 \leq B_r^2.
-		\end{equation*}
-		Therefore $\wt{u}$ satisfies~\eqref{eqn:incoherence_proof_4}. Finally, we have that
-		\begin{equation*}
-		\norm{\wt{u}}_1 = \norm{u}_1 + (u_{\star,k}  - u_k) - (\sqrt{u_l^2 - (u_{\star,k}^2 - u_k^2)} - u_l) > \norm{u}_1 + (u_{\star,k}  - u_k) - (u_{\star,k}  - u_k) = \norm{u}_1,
-		\end{equation*}
-		so we have established the desired contradiction.
-	\end{itemize}
+	Turning our attention now to the expectation $E_{\alpha,P}$, by replacing $p$ with its Taylor expansion we obtain
+	\begin{equation*}
+	E_{\alpha,P} = \int_{\Rd} (x - z)^{\alpha} K_r(x,z) p(z) \,dz \sum_{\abs{\beta} = 0}^{s - 2} \frac{p^{(\beta)}(x)}{\beta!} \underbrace{\int_{\Rd} (x - z)^{\alpha + \beta}K_r(x,z)\,dz}_{:=I_{\abs{\alpha} + \abs{\beta}}} + O(r^{s})
+	\end{equation*}
+	where the first sum equals zero when $s = 1$. Plugging this back in to~\eqref{eqn:leading_term_sobolev_pf1} yields
+	\begin{align*}
+	\Ebb(D_kf(x)) & = \sum_{\alpha = 1}^{s - 1} \sum_{\abs{\beta} = 0}^{s-2} \frac{f^{(\alpha)}(x)p^{(\beta)}(x)}{\alpha!\beta!} I_{\abs{\alpha} + \abs{\beta}} + \\ & O(r^s)\cdot\biggl(\underbrace{\sum_{\abs{\alpha} = 1}^{s-1}  f^{(\alpha)}(x) + \int_{0}^{1} \int_{B(0,1)} \sum_{\abs{\alpha} = s} f^{(\alpha)}(x + try) \,dy \,dt}_{: = g_s(x)}\biggr)
+	\end{align*}
+	where $g_s \in \Leb^2(\Rd)$ by Lemma~\ref{lem:remainder_term}.
+We are now in a position to prove the second part of~\eqref{eqn:leading_term_sobolev} where $q = 1$ and $s \in \{1,2\}$.
+When $s = 1$, the sum in the prior expression is over no terms and is equal to zero.
+Since the integral $I_{1} = 0$, the first term is also zero in the case where $s = 2$.
+For $s \in \{1,2\}$, defining $f_s := g_s$, we have $\Ebb(D_kf(x)) = O(r^s) f_s(x)$ for $f_s \in \Leb_d^2(\Xset;R)$.
+This proves the second part of~\eqref{eqn:leading_term_sobolev} when $q = 1$.
+
+	Otherwise when $s > 2$ and $q = 1$, we must analyze the first term in the prior expression.
+Since $f^{(\alpha)} \in W^{s - \abs{\alpha},2}(\Rd)$ and $p^{(\beta)} \in C^{s - 1 - \abs{\beta}}(\Rd)$, and $\min\{s - \abs{\alpha}, s - 1 - \abs{\beta}\} \geq s - (\abs{\alpha} + \abs{\beta})$, the product $f^{(\alpha)} p^{(\beta)}$ belongs to $W^{s - (\abs{\alpha} + \abs{\beta}),2}(\Rd)$, and moreover
+	\begin{equation*}
+	\norm{f^{(\alpha)} p^{(\beta)}}_{W^{s - (\abs{\alpha} + \abs{\beta}),2}(\Rd)} \leq p_{\max} \norm{f^{(\alpha)}}_{W^{s - (\abs{\alpha} + \abs{\beta}),2}(\Rd)} \leq p_{\max} \norm{f}_{W^{s,2}(\Rd)}.
+	\end{equation*} 
+	Additionally, the integral $I_1 = 0$, and for $\ell > 1$,
+	\begin{equation*}
+	\abs{I_{\ell}} \leq r^{\ell} \int K_r(x,z)\,dz = r^{\ell}
+	\end{equation*}
+	Therefore,
+	\begin{align*}
+	\sum_{\alpha = 1}^{s - 1} \sum_{\abs{\beta} = 0}^{s-2} \frac{f^{(\alpha)}(x)p^{(\beta)}(x)}{\alpha!\beta!} I_{\abs{\alpha} + \abs{\beta}} & = \sum_{\ell = 1}^{s - 1} \sum_{\substack{\abs{\alpha} + \abs{\beta} = \ell, \\ \abs{\alpha} > 0}} \frac{f^{(\alpha)}(x)p^{(\beta)}(x)}{\alpha!\beta!} I_{\ell} + O(r^{s}) \sum_{\abs{\alpha} = 2}^{s - 1} \sum_{\abs{\beta} = {s - \abs{\alpha}}}^{s - 2} \frac{f^{(\alpha)}(x)p^{(\beta)}(x)}{\alpha!\beta!} \\
+	& = \sum_{\ell = 2}^{s - 1} I_{\ell} \left( \underbrace{\sum_{\substack{\abs{\alpha} + \abs{\beta} = \ell, \\ \abs{\alpha} > 0}} \frac{f^{(\alpha)}(x)p^{(\beta)}(x)}{\alpha!\beta!}}_{:= f_{\ell}(x)} \right) + O(r^{s}) \underbrace{\sum_{\abs{\alpha} = 2}^{s} \sum_{\abs{\beta} = {s - \abs{\alpha}}}^{s - 1} \frac{f^{(\alpha)}(x)p^{(\beta)}(x)}{\alpha!\beta!}}_{:=h_{s}(x)}
+	\end{align*} 
+	and setting $f_s = g_s + h_s$ implies the first part of~\eqref{eqn:leading_term_sobolev} when $q = 1$.
+
+	\paragraph{Induction Step.}
+	In this part of the proof, the functions $f_{\ell}$ for $\ell = 2,\ldots,s$ and the constants $I_{\ell}$ for $\ell = 2,\ldots,s$ may change from line to line, but will always satisfy the conditions in the theorem statement.
+
+	Now, we assume \eqref{eqn:leading_term_sobolev} holds for all $k \in [n]^q$, and prove the desired estimate on $\Ebb(D_{kj}f(x))$ for each $j \in [n]$.
+By the law of iterated expectation and our analysis of the base case,
+	\begin{align*}
+	\Ebb(D_{kj}f(x)) & = \Ebb(D_k(\Ebb(D_jf))(x)) \\
+	& = \Ebb\left(D_k\left(\sum_{t = 2}^{s - 1} I_t f_t + O(r^s) f_s\right)(x)\right) \\
+	& = \sum_{t = 2}^{s - 1} I_t \cdot \Ebb(D_kf_t(x)) + O(r^s) f_s(x)
+	\end{align*}
+	where the second equality follows from the linearity and boundedness of $f \mapsto \Ebb(D_kf)$.
+We now apply the inductive hypothesis to $\Ebb(D_kf_t(x))$ for each $t = 2,\ldots,s-1$, to prove each part of~\eqref{eqn:leading_term_sobolev}.
+
+	First we consider the case when $2(q + 1) \geq s$.
+Note that $f_t \in W^{s - t,2}(\Rd) \cap C^{\infty}(\Rd)$, and $t \geq 2$ implies $2q \geq s - t$.
+Therefore by hypothesis $\Ebb(D_kf_t(x)) = O(r^{s - t})f_s$ for some $f_s \in \Leb^2(\Rd) \cap C^{\infty}(\Rd)$.
+As a result
+	\begin{equation*}
+	\sum_{t = 2}^{s - 1} I_t \cdot \Ebb(D_kf_t(x)) = \sum_{t = 2}^{s - 1} I_t \cdot O(r^{s - t}) f_s(x) = O(r^s) f_s(x)
+	\end{equation*}
+	establishing that the second part of~\eqref{eqn:leading_term_sobolev} holds for all $q$.
+
+	Otherwise $2(q + 1) < s - 1$.
+For each $t = 2,\ldots, s - 1$, if additionally  $2q \leq s - t - 1$, then by hypothesis $\Ebb(D_kf_t(x)) = \sum_{\ell = 2q}^{s - t - 1} I_{\ell} \cdot f_{\ell + t}(x) + O(r^{s - t}) f_s$, and otherwise $\Ebb(D_kf_t(x)) = O(r^{s - t}) f_s(x)$.
+Therefore,
+	\begin{align*}
+	\sum_{t = 2}^{s - 1} I_t \cdot \Ebb(D_kf_t(x)) & = \sum_{t = 2}^{s - 1 - 2q} I_{t} \cdot \left\{\sum_{\ell = 2q}^{s - t - 1} I_{\ell} \cdot f_{\ell + t}(x) + O(r^{s - t}) \cdot f_s(x) \right\} + \sum_{t = s - 1 - 2q}^{s - 1}I_{t} \cdot O(r^{s - t}) \cdot f_s(x) \\
+	& = \sum_{t = 2}^{s - 1 - 2q} I_t \cdot \left\{\sum_{\ell = 2q}^{s - t - 1} I_{\ell} \cdot f_{\ell + t}(x)\right\}  + O(r^s)\cdot f_s(x) \\
+	& = \sum_{\ell = 2q}^{s - 3} \sum_{t = 2}^{s - \ell - 1} I_{\ell + t} \cdot f_{\ell + t}(x) + O(r^s)\cdot f_s(x).
+	\end{align*}
+	Rewriting the final equation as a sum over $\ell + t = 2(q + 1),\ldots, s - 1$ establishes~\eqref{eqn:leading_term_sobolev}.
 \end{proof}
-
-\paragraph{Putting the pieces together.}
-We apply Lemma~\ref{lem:pi_max_pf_1_util_1} to the vector $u = (u_i)_{i = 1}^{n}$, where
-\begin{equation*}
-u_i = \frac{1}{\sqrt{2}}\left(\abs{\dotp{v_k}{\overline{v}_{k - i}}} + \abs{\dotp{v_k}{\overline{v}_{k + i}}}\right).
-\end{equation*}
-Note the following:
-\begin{itemize}
-	\item $\norm{u}_2 \leq 1$.
-	\item By \eqref{eqn:incoherence_proof_2}, the vector $u$ satisfies the constraint~\eqref{eqn:incoherence_proof_4} with
-	\begin{equation*}
-	B_r^2 = 400\frac{n^4\delta^2s_{n}^2}{r^4} ~~\textrm{when $r \geq 8 k \sqrt{\delta}$.}
-	\end{equation*}
-	\item Taylor expanding $f(r + 1) = (r + 1)^{-4}$ around $r$, we have that
-	\begin{equation*}
-	B_r^2 - B_{r + 1}^2 \leq 3200 \frac{n^4 \delta^2 s_{n}^2}{r^5} ~~\textrm{when $r \geq 2$.}
-	\end{equation*}
-	Therefore, the sequence $(B_r)$ satisfies \eqref{eqn:incoherence_util_1} for all $r$ large enough such that 
-	\begin{equation*}
-	r \geq 2^{5/4} 100^{1/4} n \sqrt{\delta s_{n}}
-	\end{equation*}
-	\item Since $u$ and $(B_r)$ satisfy the constraints \eqref{eqn:incoherence_proof_3}, \eqref{eqn:incoherence_proof_4} and \eqref{eqn:incoherence_proof_5}, we may apply Lemma~\ref{lem:pi_max_pf_1_util_1} to $u$, obtaining that
-	\begin{equation*}
-	\sum_{j = 1}^{n} \abs{\dotp{v_j}{\overline{v}_k}} = \sqrt{2}\norm{u}_1 \leq \sqrt{2R} + 60\sqrt{2} n^2 \delta s_{n} \sum_{r = R}^{n} r^{-5/2}
-	\end{equation*}
-	for any $R \geq 2 \vee k\sqrt{\delta} \vee 2^{5/4} 100^{1/4} n \sqrt{\delta s_{n}}$. Bounding sum by integral, we have that if $R \geq 2$ then
-	\begin{equation*}
-	\sum_{r = R}^{n} r^{-5/2} \leq \int_{R - 1}^{n - 1} x^{-5/2} \,dx \leq \frac{2^{3/2}2}{3} R^{-3/2}.
-	\end{equation*}
-	Plugging this into the previous expression, we arrive at
-	\begin{equation*}
-	\sum_{j = 1}^{n} \abs{\dotp{\overline{v}_j}{v_k}} \leq \sqrt{R} + 60 \sqrt{2} n^2 \delta s_{\max} \frac{2^{3/2}2}{3R^{3/2}}
-	\end{equation*}
-	for any $R \geq 2 \vee k\sqrt{\delta} \vee 2^{5/4} 100^{1/4} n \sqrt{\delta s_{\max}}$. Lemma~\ref{lem:pi_max_pf_1} then follows from~\eqref{eqn:incoherence_proof_1}.
-\end{itemize}
-
-\section{Auxiliary Results}
-
-\subsection{Results of Others}
-
-Let $\overline{X}$ denote the $n$ evenly spaced grid points on $[0,1]^d$; formally, letting $t = n^{1/d}$
-\begin{equation*}
-\overline{X} = \biggl\{\frac{1}{t}(k_1,\ldots,k_d): k \in [t]^d\biggr\}
-\end{equation*}
-
-\begin{theorem}[Theorem 1 of \textcolor{red}{Garcia-Trillos and Slepcev}]
-	\label{thm:slepcev_transport_distance}
-	With probability at least $1 - c/n$ there exists a bijection $\pi: \overline{X} \to X$ such that
-	\begin{equation}
-	\label{eqn:slepcev_transport_distance}
-	\max_{k \in [t]^d} \abs{\overline{x}_k - \pi(\overline{x}_k)} \leq c \left(\frac{\log n}{n}\right)^{1/d}
-	\end{equation}
-\end{theorem}
-
-\begin{theorem}[\textcolor{red}{Evans} Chapter 5.4, Theorem 1]
-	\label{thm:evans_extension}
-	Assume $U$ is bounded and $\partial U$ is $C^1$.
-Select a bounded open set $V$ such that $U \subset \subset V$ ($U$ is compactly contained in $V$).
-Then there exists a bounded linear operator $E: W^{1,2}(U) \to W^{1,2}(\Rd)$ such that for each $u \in W^{1,2}(U)$:
-	\begin{enumerate}
-		\item $Eu = u$ a.e. in $U$,
-		\item $Eu$ has support within $V$, and 
-		\item 
-		\begin{equation*}
-		\norm{Eu}_{W^{1,2}(\Rd)} \leq C \norm{u}_{W^{1,2}(\Rd)}
-		\end{equation*}
-		the constant $C$ depending only on $U$ and $V$.
-	\end{enumerate}
-\end{theorem}
-
-\begin{lemma}[Poincare inequality for path graphs.]
-	\label{lem: path_poincare}
-	Fix $m \geq 0$. For vertices $V = \set{1, \ldots,m}$ define the path $P(1 \to m) = ((1,2),(2,3),\ldots, (m-1,m))$ and $G_{(1,m)}$ to be the graph consisting only of an edge between $1$ and $m$. Then,
-	\begin{equation*}
-	(m - 1) \cdot P(1 \to m) \succeq G_{(1,m)}
-	\end{equation*}
-\end{lemma}
-
-\subsection{Integrals}
 
@@
 	\label{lem:remainder_term}
-	Suppose $f \in \Leb^2(\Rd)$, and $h:\Rd \times \Rd \times [0,1] \to \Reals$ is uniformly bounded.
-Then, the function $g(x) = \int_{0}^{1} \int_{B(0,1)} f(x + aty) h(y,x, t) \,dy \,dt$ also belongs to $\Leb^2(\Rd)$, with norm
-	\begin{equation*}
-	\norm{g}_{\Leb^2(\Rd)} \leq \nu_d \cdot \norm{f}_{\Leb^2(\Rd)} \cdot \norm{h}_{\infty}
+	Suppose $f \in \Leb^2(\Rd)$. Then, the function $g(x) = \int_{0}^{1} \int_{B(0,1)} f(x + aty) \,dy \,dt$ also belongs to $\Leb^2(\Rd)$, with norm
+	\begin{equation*}
+	\norm{g}_{\Leb^2(\Rd)} \leq \nu_d\cdot \norm{f}_{\Leb^2(\Rd)}
 	\end{equation*}
@@
 	\begin{align*}
-	\norm{g}_{\Leb^2(\Rd)}^2 & = \int_{\Rd} \left(\int_{0}^{1} \int_{B(0,1)} f(x + aty) h(y,x,t) \,dt \,dy \right)^2 \,dx \\
-	& \leq \norm{h}_{\infty}^2 \int_{\Rd} \left(\int_{0}^{1} \int_{B(0,1)} f(x + aty) \,dt \,dy \right)^2 \,dx \\
-	& \leq \nu_d^2 \norm{h}_{\infty}^2 \int_{\Rd} \int_{0}^{1} \frac{1}{\nu_d}\int_{B(0,1)} f^2(x + aty) \,dt \,dy \,dx \tag{Jensen's inequality} \\
-	& = \nu_d^2 \norm{h}_{\infty}^2 \int_{0}^{1} \int_{B(0,1)} \frac{1}{\nu_d}\int_{\Rd}f^2(x + aty) \,dt \,dy \,dx \tag{Fubini's theorem} \\
-	& = \nu_d^2 \norm{h}_{\infty}^2 \norm{f}_{\Leb^2(\Rd)}^2.
+	\norm{g}_{\Leb^2(\Rd)}^2 & = \int_{\Rd} \left(\int_{0}^{1} \int_{B(0,1)} f(x + aty) \,dt \,dy \right)^2 \,dx \\
+	& \leq \nu_d^2 \int_{\Rd} \int_{0}^{1} \frac{1}{\nu_d}\int_{B(0,1)} f^2(x + aty) \,dt \,dy \,dx \tag{Jensen's inequality} \\
+	& = \nu_d^2  \int_{0}^{1} \int_{B(0,1)} \frac{1}{\nu_d}\int_{\Rd}f^2(x + aty) \,dt \,dy \,dx \tag{Fubini's theorem} \\
+	& = \nu_d^2 \norm{f}_{\Leb^2(\Rd)}^2.
 	\end{align*}
@@
 
-\begin{lemma}
-	\label{lem:leading_term_sobolev}
-	Let $k \in [n]^q$ for some $q \geq 1$, and let $K_r$ be a second order kernel.
-For $s \geq 0$ an integer, let $p$ be a density satisfying $p \in C^{s - 1}(\Rd;p_{\max})$ for some $p_{\max} > 0$ if $s \geq 1$, and otherwise $p \in C^{0}(\Rd;p_{\max})$.
-Then for any $f \in C^{\infty}(\Rd)$, there exist functions $f_{\ell}, \ell = 2q,\ldots,s$ satisfying $f_{\ell} \in C^{\infty}(\Rd)$ and
-	\begin{equation*}
-	\norm{f_{\ell}}_{W^{s - \ell,2}(\Rd)} \leq c \norm{f}_{W^{s,2}(\Rd)}
-	\end{equation*}
-	for some constant $c$ which depends only on $d$, $\Xset$ and $p_{\max}$ such that
-	\begin{equation}
-	\label{eqn:leading_term_sobolev}
-	\Ebb(D_kf(x)) =
+\section{Proof of~\eqref{eqn:higher_order_sobolev_testing_rate_pf2}}
+\begin{lemma}
+	\label{lem:empirical_norm_sobolev}
+	Let $\Xset$ be a Lipschitz domain over which the density is upper and lower bounded 
+	\begin{equation*}
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty ~~\textrm{for all $x \in \Xset$,}
+	\end{equation*}
+	and let $f \in W_d^{s,2}(\Xset)$.Then for any $b \geq 1$, there exists $c_1$ such that if 
+	\begin{equation}
+	\norm{f}_{\Leb^2(\Xset)} \geq 
 	\begin{cases*}
-	\sum_{\ell = 2q}^{s} f_{\ell}(x) r^{\ell} ,~~& \textrm{if $2q < s$} \\
-	r^s \cdot f_{s}(x),~~& \textrm{if $2q \geq s$}
+	c_1 \cdot b \cdot \norm{f}_{W_d^{s,2}(\Xset)} \cdot \max\{n^{-1/2},n^{-s/d}\},~~\textrm{if}~2s \neq d \\
+	c_1 \cdot b \cdot \norm{f}_{W_d^{s,2}(\Xset)} \cdot n^{-a/2},~~\textrm{if}~ 2s = d ~\textrm{for any}~ 0 < a < 1
 	\end{cases*}
 	\end{equation}
+	then,
+	\begin{equation}
+	\label{eqn:l2_to_empirical_norm_sobolev}
+	\mathbb{P}\left[\norm{f}_n^2 \geq \frac{1}{b} \Ebb\bigl[\norm{f}_n^2\bigr]\right] \geq 1 - \frac{5}{b}
+	\end{equation}
+	where $c_1$ and $c_2$ are constants which may depend only on $s$, $\Xset$, $d$, $p_{\min}$ and $p_{\max}$.
 \end{lemma}
 \begin{proof}
-	We proceed by induction on $q$.
-Note that without loss of generality, we will assume $p \in C^{\infty}(\Rd)$, i.e.
-$p$ is smooth.
-Otherwise we can simply mollify $p$, with neglible change to the Holder norm $\norm{p}_{C^{s-1}(\Xset)}$ or the expectation $\Ebb[D_kf(x)]$.
-
-	\paragraph{Base case.}
-	We begin with the base case of $q = 1$. When $s = 0$, we need to show that $\Ebb[D_kf(x)] = f_s(x)$ for some $\norm{f_s}_{\Leb^2(\Rd)} \leq c\norm{f}_{W^{s,2}(\Rd)}$. But
-	\begin{align*}
-	\Ebb\Bigl[D_kf(x)\Bigr] & = \int f(z) K_r(x,z) p(z) \,dz  - f(x) \\
-	& = \int f(yr + x) K(\norm{y}) p(yr + z) \,dy - f(x).
-	\end{align*}
-	and since
-	\begin{equation*}
-	\biggl\|\int f(yr + x) K(\norm{y}) p(yr + x) \,dy\biggr\|_{\Leb^2(\Rd)} \leq p_{\max} K_{\max} \norm{f}_{\Leb^2(\Rd)}
-	\end{equation*}
-	the claim follows.
-
-	Now, when $s \geq 1$ since $f$ is smooth it admits a Taylor expansion of the following form for all $x,z \in \Rd$:
-	\begin{equation*}
-	f(z) = \sum_{\abs{\alpha} < s} \frac{f^{(\alpha)}(x)}{\alpha!} (x - z)^{\alpha} + \frac{1}{(s - 1)!}\sum_{\abs{\alpha} = s} (x - z)^{\alpha} \int_{0}^{1}(1 - t)^{s - 1} f^{(\alpha)}(x + t(z - x)) \,dt
-	\end{equation*}
-	where $f^{(\alpha)} \in C^{\infty}(\Rd)$ additionally satisfies
-	\begin{equation*}
-	\norm{f^{(\alpha)}}_{W^{s - \abs{\alpha},2}(\Rd)} \leq \norm{f}_{W^{s,2}(\Rd)}
-	\end{equation*}
-	Replacing $f$ by its Taylor expansion inside the expected first order difference operator $\Ebb(D_kf(x))$ we have
-	\begin{equation}
-	\label{eqn:leading_term_sobolev_pf1}
-	\Ebb(D_kf(x)) = \sum_{1 < \abs{\alpha} < s} \frac{f^{(\alpha)}(x)}{\alpha!} E_{\alpha,x} + \frac{1}{(s - 1)!}\sum_{\abs{\alpha} = s} \int_{0}^{1}(1 - t)^{s - 1} E_{\alpha,x,t}(f)  \,dt 
-	\end{equation}
-	where we use the notation $E_{\alpha,x} := \Ebb\left[(x - x_k)^{\alpha}K_r(x_k,x)\right]$ and  $E_{\alpha,x,t}(f) := \Ebb\bigl[f^{(\alpha)}(x + t(x_k - x)) (x_k - x)^{\alpha} K_r(x_k,x)\bigr]$.
-
-	 By a change of variables, we have
-	\begin{align*}
-	E_{\alpha,x,t}(f) & = \frac{1}{r^d}\int_{\Rd} f^{(\alpha)}\bigl(x + t(z - x)\bigr) (z - x)^{\alpha}  K\Bigl(\frac{\norm{z - x}}{r}\Bigr) p(z) \,dz \\
-	& = r^s \int_{\Rd} y^{\alpha} f^{(\alpha)}\bigl(x + tyr) K(\norm{y}) p(yr + x) \,dy,
-	\end{align*}
-	and as a result the remainder term in \eqref{eqn:leading_term_sobolev_pf1} reduces to
-	\begin{align*}
-	\frac{1}{(s - 1)!}\sum_{\abs{\alpha} = s} \int_{0}^{1}(1 - t)^{s - 1} E_{\alpha,x,t}(f)  \,dt & = \frac{r^s}{(s - 1)!}\sum_{\abs{\alpha} = s} \int_{0}^{1} \int_{\Rd} (1 - t)^{s - 1}  y^{\alpha} f^{(\alpha)}\bigl(x + tyr) K(\norm{y}) p(yr + x) \,dy \,dt \\
-	& =: r^s g_s(x) ~~\textrm{for $g_s \in C^{\infty}(\Rd)$} 
-	\end{align*}
-	where $\norm{g_s}_{\Leb^2(\Rd)} \leq c \norm{f}_{W^{s,2}(\Rd)}$ follows from Lemma~\ref{lem:remainder_term}.
-When $s = 1$, there are no $1 < \abs{\alpha} < s$, and so only the remainder term in \eqref{eqn:leading_term_sobolev_pf1} is non-zero; thus we have shown~\eqref{eqn:leading_term_sobolev} when $q = 1$ and $s = 1$.
-
-	When $s \geq 2$, we can analyze $E_{\alpha,x}$ using a Taylor expansion of $p$. For any $x,z \in \Rd$, we have
-	\begin{equation*}
-	p(z) = \sum_{\abs{\beta} < s  - 1} \frac{p^{(\beta)}(x)}{\beta!} (x - z)^{\beta} + \frac{1}{(s - 2)!}\sum_{\abs{\beta} = s - 1} (x - z)^{\beta} \int_{0}^{1}(1 - t)^{s - 2} p^{(\beta)}(x + t(z - x)) \,dt
-	\end{equation*}
-	and $p^{(\beta)} \in C^{\infty}(\Rd)$ additionally satisfies
-	\begin{equation*}
-	\norm{p^{(\beta)}}_{C^{s - \abs{\beta} - 1}(\Rd)} \leq p_{\max}.
-	\end{equation*}
-	Replacing $p$ by its Taylor expansion, the term $E_{\alpha,x}$ can be rewritten as
-	\begin{align*}
-	E_{\alpha,x} & = \int_{\Rd} (x - z)^{\alpha} K_r(x,z) p(z) \,dz \\
-	& =  \sum_{\abs{\beta} = 0}^{s - 2} \frac{p^{(\beta)}(x)}{\beta!} \int_{\Rd} (x - z)^{\alpha + \beta}K_r(x,z)\,dz ~~+ \\
-	& ~~~\frac{1}{(s - 2)!}\sum_{\abs{\beta} = s - 1} \int_{\Rd} \int_{0}^{1} (1 - t)^{s - 2} (x - z)^{\alpha + \beta - 1}  p^{(\beta)}(x + t(z - x)) K_r(x,z) \,dz \,dt \\
-	& = \sum_{\abs{\beta} = 0}^{s - 2} r^{\abs{\alpha} + \abs{\beta}} \frac{p^{(\beta)}(x)}{\beta!} \int_{\Rd} y^{\alpha + \beta}K(y)\,dy +~ \frac{r^{\abs{\alpha} + s - 1}}{(s - 2)!} \sum_{\abs{\beta} = s - 1} \int_{\Rd} \int_{0}^{1} (1 - t)^{s - 2} y^{\alpha + \beta - 1} p^{(\beta)}(x + ytr) K(\norm{y}) \,dy \,dt \\
-	& =: \sum_{\abs{\beta} = 0}^{s - 2} \1\{\abs{\alpha} + \abs{\beta} > 1\}r^{\abs{\alpha} + \abs{\beta}} p_{\abs{\beta}}(x) + r^s p_s(x) ~~\textrm{for $p_{\abs{\beta}} \in C^{\infty}(\Rd), \norm{p_{\abs{\beta}}}_{C^{s - \abs{\beta} - ,1}(\Rd)} \leq c \norm{p}_{C^{s - 1}(\Rd)}$},
-	\end{align*}
-	where the last line follows since $K$ is a 2nd order kernel. 
-	Substituting our expressions for $E_{\alpha,x}$ and $E_{\alpha,x,t}(f)$ back into~\eqref{eqn:leading_term_sobolev_pf1}, we obtain
-	\begin{align*}
-	\Ebb\bigl[D_kf(x)\big] & = \sum_{\abs{\alpha} = 1}^{s - 1} \sum_{\abs{\beta} = 0}^{s-2} \1\{\abs{\alpha} + \abs{\beta} > 1\} r^{\abs{\alpha} + \abs{\beta}} \frac{f^{(\alpha)}(x)p_{\abs{\beta}}(x)}{\alpha!} + r^s \biggl(\sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)p_s(x)}{\alpha!} + g_s(x)\biggr) \\
-	& = \sum_{\ell = 2}^{s - 1} r^{\ell} \sum_{ \substack{\abs{\alpha} + \abs{\beta} = \ell, \\ \abs{\beta} > 0} }   \frac{f^{(\alpha)}(x)p_{\abs{\beta}}(x)}{\alpha!} + r^s \biggl(\sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)p_s(x)}{\alpha!} + g_s(x)\biggr) \\
-	& =: \sum_{\ell = 2}^{s - 1} r^{\ell} f_{\ell}(x) + r^s f_s(x).
-	\end{align*}
-	It remains to show that $\norm{f_{\ell}}_{W^{s - \ell,2}(\Rd)} \leq c \norm{f}_{W^{s,2}(\Rd)}$ for each $\ell = 2,\ldots,s$.
-To see this, note that for $\abs{\alpha} + \abs{\beta} = \ell$ and $\min\{s - \abs{\alpha}, s - 1 - \abs{\beta}\} = m$, the inequality $s - \ell \leq m$ holds.
+	To prove~\eqref{eqn:l2_to_empirical_norm_sobolev} we will show
+	\begin{equation*}
+	\mathbb{E}\bigl[\norm{f}_n^4\bigr] \leq \left(1 + \frac{1}{b^2}\right)\cdot \left(\mathbb{E}\bigl[\norm{f}_n^2\bigr]\right)^2
+	\end{equation*}
+	whence the claim follows from the Paley-Zygmund inequality (Lemma~\ref{lem:paley_zygmund}).
+Since $p \leq p_{\max}$ is uniformly bounded, we can relate $\mathbb{E}\bigl[\norm{f}_n^4\bigr]$ to the $\Leb^4$ norm,
+	\begin{equation*}
+	\mathbb{E}\bigl[\norm{f}_n^4\bigr] = \frac{(n-1)}{n}\left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + \frac{\mathbb{E}\Bigl[\bigl(f(x_1)\bigr)^4\Bigr]}{n} \leq \left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + p_{\max}^2\frac{\norm{f}_{\Leb^4}^4}{n}.
+	\end{equation*}
+	We will use a Sobolev inequality to relate $\norm{f}_{\Leb^4}$ to $\norm{f}_{W_d^{s,2}(\Xset)}$.
+The nature of this inequality depends on the relationship between $s$ and $d$ (see Theorem 6 in Section 5.6.3 of \textcolor{red}{Evans} for a formal statement), so from this point on we divide our analysis into three cases: (i) the case where $2s > d$, (ii) the case where $2s < d$, and (iii) the borderline case $2s = d$.
+
+	\paragraph{Case 1: $2s > d$.}
+	When $2s > d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in C^{\gamma}(\overline{\Xset})$ for some $\gamma > 0$ which depends on $s$ and $d$, with the accompanying estimate
+	\begin{equation*}
+	\sup_{x \in \Xset} \abs{f(x)} \leq \norm{f}_{C^{\gamma}(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}.
+	\end{equation*}
 	Therefore,
-	\begin{equation*}
-	\norm{f^{(\alpha)}p_{\abs{\beta}}}_{W^{s - \ell,2}(\Rd)} \leq \norm{f^{(\alpha)}p_{\abs{\beta}}}_{W^{m,2}(\Rd)} \leq p_{\max} \norm{f^{(\alpha)}}_{W^{s - \abs{\alpha},2}(\Xset)} \leq p_{\max} \norm{f}_{W^{s,2}(\Rd)}
-	\end{equation*} 
-	which finishes the proof of Lemma~\ref{lem:leading_term_sobolev} when $q = 1$.
-
-	\paragraph{Induction Step.}
-
-	Now, we assume \eqref{eqn:leading_term_sobolev} holds for all $k \in [n]^q$, and prove the desired estimate on $\Ebb[D_{k}D_jf(x)]$ for each $j \in [n]$.
-We will build a proof piece-by-piece, depending on the relative size of $s$ and $q$.
-
-	If $s \leq 2$, by hypothesis there exists $f_s$ satisfying
-	\begin{equation*}
-	\norm{f_s}_{\Leb^2(\Rd)} \leq c \norm{f}_{W^{s,2}(\Rd)}
-	\end{equation*}
-	such that for any $z \in \Rd$,
-	\begin{equation*}
-	\Ebb\Bigl[D_jf(z)\Bigr] = r^s f_s(z).
-	\end{equation*}
-	Therefore by the law of iterated expectation,
-	\begin{equation*}
-	\Ebb\Bigl[D_kD_jf(x)\Bigr] = \Ebb\Bigl[D_k(\Ebb[D_jf])(x)\Bigr] = \Ebb\Bigl[D_k\bigl(r^s f_s\bigr)(x)\Bigr] = r^s \Ebb\Bigl[D_kf_s(x)\Bigr]
-	\end{equation*}
-	and again using the inductive hypothesis, we have that
-	\begin{equation}
-	\label{eqn:leading_term_sobolev_pf3}
-	\biggl\|\Ebb\Bigl[D_kf_s(x)\Bigr]\biggr\|_{\Leb^2(\Rd)} \leq c \norm{f_s}_{\Leb^2(\Rd)},
-	\end{equation}
-	establishing~\eqref{eqn:leading_term_sobolev} for all $q$ in the case when $s \leq 2$.
-
-	Otherwise, when $s \geq 3$ by hypothesis there exist functions $f_{\ell}$ for $\ell = 2,\ldots,s$ satisfying
-	\begin{equation*}
-	\norm{f_{\ell}}_{W^{s - \ell,2}(\Rd)} \leq c \norm{f}_{W^{s,2}(\Rd)}
-	\end{equation*}
-	such that for any $z \in \Rd$,
-	\begin{equation*}
-	\Ebb(D_jf(z)) = \sum_{\ell = 2}^{s} r^{\ell} f_{\ell}(z).
-	\end{equation*}
-	As a result, by the law of iterated expectation
-	\begin{align*}
-	\Ebb\Bigl[D_{k}D_jf(x)\Bigr] & = \Ebb\Bigl[D_k\bigl(\Ebb[D_jf]\bigr)(x)\Bigr] \\
-	& = \Ebb\biggl[D_k\Bigl(\sum_{\ell = 2}^{s - 1} r^{\ell} f_{\ell} + r^s f_s\Bigr)(x)\biggr] \\
-	& = \sum_{\ell = 2}^{s - 1} r^{\ell} \cdot \Ebb\Bigl[D_kf_{\ell}(x)\Bigr] + r^s \Ebb\Bigl[D_kf_s(x)\Bigr].
-	\end{align*}
-	Recalling that $\Ebb\Bigl[D_kf_s(x)\Bigr]$ satisfies~\eqref{eqn:leading_term_sobolev_pf3}, we now apply the inductive hypothesis to $\Ebb(D_kf_{\ell}(x))$ for each $\ell = 2,\ldots,s-1$, to prove~\eqref{eqn:leading_term_sobolev}.
-
-	First we consider the case when $2(q + 1) \geq s$.
-Note that $\ell \geq 2$ implies $2q \geq s - \ell$.
-Therefore by hypothesis, for each $\ell = 2,\ldots,s - 1$ the expectation $\Ebb[D_kf_{\ell}(x)] = r^{s - \ell}f_{\ell,s}(x)$ for some $f_{\ell,s} \in \Leb^2(\Rd)$ satisfying
-	\begin{equation}
-	\label{eqn:leading_term_sobolev_pf4}
-	\norm{f_{\ell,s}}_{\Leb^2(\Rd)} \leq c\norm{f_{\ell}}_{W^{s - \ell,2}(\Rd)} \leq c\norm{f}_{W^{s,2}(\Rd)}
-	\end{equation}
-	and as a result
-	\begin{equation*}
-	\sum_{\ell = 2}^{s - 1} r^{\ell} \cdot \Ebb(D_kf_{\ell}(x)) = r^s \sum_{\ell = 2}^{s - 1} f_{\ell,s}(x)
-	\end{equation*}
-	establishing that the second part of~\eqref{eqn:leading_term_sobolev} holds for all $q$ and all $s \leq 2(q + 1)$.
-
-	Otherwise $2(q + 1) < s - 1$.
-For each $\ell = 2,\ldots, s - 1$, if additionally  $2q \leq s - \ell - 1$, then by hypothesis $\Ebb(D_kf_{\ell}(x)) = \sum_{m = 2q}^{s - \ell - 1} r^{m} \cdot f_{\ell, \ell + m}(x) + r^{s - {\ell}} f_{\ell,s}$, and otherwise $\Ebb(D_kf_{\ell}(x)) = r^{s - \ell}f_{\ell,s}(x)$, where for each $\ell$ and $m$,
-	\begin{equation*}
-	\norm{f_{\ell,\ell + m}}_{W^{s - (\ell + m),2}(\Rd)} \leq c\norm{f_{\ell}}_{W^{s - \ell}} \leq c\norm{f}_{W^{s,2}(\Rd)}
-	\end{equation*}
-	and $f_{\ell,s}$ satisfies \eqref{eqn:leading_term_sobolev_pf4}. Therefore,
-	\begin{align*}
-	\sum_{\ell = 2}^{s - 1}  r^{\ell} \cdot \Ebb(D_kf_\ell(x)) & = \sum_{\ell = 2}^{s - 1 - 2q} r^{\ell} \cdot \left\{\sum_{m = 2q}^{s - \ell - 1} r^m \cdot f_{\ell,\ell + m}(x) + r^{s - \ell} \cdot f_{\ell,s}(x) \right\} + \sum_{\ell = s - 1 - 2q}^{s - 1} r^s \cdot f_{\ell,s}(x) \\
-	& = \sum_{\ell = 2}^{s - 1 - 2q} r^{\ell} \cdot \left\{\sum_{m = 2q}^{s - \ell - 1} r^m \cdot f_{\ell, \ell + m}(x)\right\}  + r^{s}\sum_{\ell = 2}^{s - 1}f_{\ell,s}(x) \\
-	& = \sum_{m = 2q}^{s - 3} \sum_{\ell = 2}^{s - m - 1} r^{m + \ell} \cdot f_{\ell, \ell + m}(x) + r^{s}\sum_{\ell = 2}^{s - 1}f_{\ell,s}(x).
-	\end{align*}
-	Rewriting the first sum in the final equation as a sum over $\ell + m = 2(q + 1),\ldots, s - 1$ establishes~\eqref{eqn:leading_term_sobolev}.
+	\begin{align*}
+	\norm{f}_{\Leb^4}^4 & = \int_{\Xset} [f(x)]^4 \,dx \\
+	& \leq \left(\sup_{x \in \Xset} \abs{f(x)}\right)^2 \cdot \int_{\Xset} [f(x)]^2 \,dx \\
+	& \leq c \norm{f}_{W^{s,2}(\Xset)}^2 \cdot \norm{f}_{\Leb^2(\Xset)}^2.
+	\end{align*}
+	Since by assumption
+	\begin{equation*}
+	\norm{f}_{\Leb^2(\Xset)}^2 \geq c_1^2 \cdot b^2 \cdot \norm{f}_{W_d^{s,2}(\Xset)}^2 \cdot \frac{1}{n},
+	\end{equation*}
+	we have
+	\begin{equation*}
+	p_{\max}^2\frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c \norm{f}_{W^{s,2}(\Xset)}^2 \cdot \frac{\norm{f}_{\Leb^2(\Xset)}^4}{n \norm{f}_{\Leb^2(\Xset)}^2} \leq c \frac{\norm{f}_{\Leb^2(\Xset)}^4}{c_1^2 b^2} \leq \frac{\Ebb\bigl[\norm{f}_n^2\bigr]}{b^2},
+	\end{equation*}
+	where the last inequality follows by taking $c_1$ sufficiently large.
+
+	\paragraph{Case 2: $2s < d$.}
+	When $2s < d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in \Leb^q(\Xset)$ for $q = 2d/(d - 2s)$, and moreover that
+	\begin{equation*}
+	\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}.
+	\end{equation*}
+	Since $4 = 2\theta + (1 - \theta)q$ for $\theta = 2 - d/(2s)$, Lyapunov's inequality implies
+	\begin{equation*}
+	\norm{f}_{\Leb^4(\Xset)}^4 \leq \norm{f}_{\Leb^2}^{2\theta} \cdot \norm{f}_{\Leb^q(\Xset)}^{(1 - \theta)q} \leq c \norm{f}_{\Leb^2(\Xset)}^{4} \cdot \left(\frac{\norm{f}_{W^{s,2}(\Xset)}}{\norm{f}_{\Leb^2(\Xset)}}\right)^{d/s}.
+	\end{equation*}
+	By assumption, $\norm{f}_{\Leb^2(\Xset)} \geq c_1 b \norm{f}_{W^{s,2}(\Xset)} n^{-s/d}$, and therefore
+	\begin{equation*}
+	p_{\max}^2 \frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c\norm{f}_{\Leb^2(\Xset)}^4 \left(\frac{\norm{f}_{W^{s,2}(\Xset)}}{n^{s/d}\norm{f}_{\Leb^2(\Xset)}}\right)^{d/s} \leq \frac{c\norm{f}_{\Leb^2(\Xset)}^4}{c_1b^{d/s}} \leq \frac{\norm{f}_{\Leb^2(\Xset)}^4}{b^2}.
+	\end{equation*}
+	where the last inequality follows when $c_1$ is sufficiently large, and keeping in mind that $d/s > 2$ and $b \geq 1$.
+
+	\paragraph{Case 3: $2s = d$.}
+	Assume $f$ satisfies~\eqref{eqn:paley_zygmund_1} for a given $0 < a < 1$.
+When $2s = d$, since $\Xset$ is a Lipschitz domain we have that $f \in L^q(\Xset)$ for any $q < \infty$, with the accompanying estimate
+	\begin{equation*}
+	\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}.
+	\end{equation*}
+	In particular the above holds for $q = 2/(1 - a)$ when $1/2 < a < 1$, and for any $q > 4$ when $0 < a < 1/2$. Using Lyapunov's inequality as in the previous case then implies the desired result.
 \end{proof}

### 2020-01-25 notes/weekly_notes/1.3.20.NotesForWeek.tex
updated weekly notes and project level notes

@@
 Suppose we observe independent samples $X = \set{x_1,\ldots,x_n}$ from a distribution $P$ with density $p$ supported on $\mathcal{X} = [0,1]^d$.
-Let $G_{n,r}$ be the random geometric graph of radius $r$ formed over the samples $X$, that is, the neighborhood graph with kernel $K(x) = \1(\norm{x} \leq r)$, and let $L$ denote the corresponding graph Laplacian matrix.
+For a given kernel $K_r(x,y) = \frac{1}{r^d}K(\norm{x - y})$, let $G_{n,r}$ denote the random neighborhood graph formed over the samples $X$.
+Let $L = D - A$ be the corresponding combinatorial Laplacian.
 In this week's notes, we will show that as long as $r$ is chosen to be sufficiently small, the eigenvalues of $L$ we care about scale at an appropriate rate.
-More precisely, we will prove the following Lemma.
+
+\section{Random Geometric Graph}
+In the special case when $K(x) = 1(x \leq r)$ and $G_{n,r}$ is a random geometric graph, we can use results from optimal transport theory to  will prove the following Lemma.
 \begin{lemma}
@@
 Along with an upper bound on the graph Sobolev semi-norm $R_{s,n}(f)$ (established in the \textit{12.20.19 Notes}), this allows us to argue that if $f$ is a smooth function with large $\mathcal{L}_2$ norm, then the projection of a function $f$ onto the eigenvectors $v_1,\ldots,v_{\kappa}$ is still quite large (measured in empirical norm); or, equivalently, the bias induced by projection onto these eigenvectors is small.
+
+\section{Higher-order kernels.}
+When $K$ is the uniform kernel, the aforementioned upper bound on the graph Sobolev semi-norm holds only when $s = 1$ or $s = 2$.
+For larger values of $s$, we require that $K$ be a higher-order kernel, that is that $K$ satisfy
+\begin{equation*}
+\int K(x) \,dx = 1,~~ \int x^{\ell}K(x) \,dx = 0~~\textrm{for $\ell = 1,\ldots,s-1$}
+\end{equation*}
+It seems difficult to extend the techniques used to prove Lemma~\ref{lem:higher_order_eigenvalue_tail_bound} to this setting. We will therefore use a different approach to prove the following Lemma.
+\begin{lemma}
+	Let $\kappa = n^{2d/(4s + d)}$. Suppose $p$ is uniformly bounded away from $0$ and $\infty$ on its support,
+	\begin{equation*}
+	0 < p_{\min} < p(x) < p_{\max} < \infty,~~\textrm{for all $x \in \Xset$.}
+	\end{equation*}
+	Let $K$ satisfy the following conditions:
+
+	\textcolor{red}{TODO:} Specify conditions.
+
+	Then there exists a constant $c$ such that the eigenvalue $\lambda_{\kappa}$ is lower bounded,
+	\begin{equation}
+	\label{eqn:higher_order_eigenvalue_tail_bound2}
+	\lambda_{\kappa} \geq c\kappa^{2/d}nr^{2 + d}
+	\end{equation}
+	with probability $1 - o(n^{-1})$ as $n \to \infty$.
+\end{lemma}
 
@@
 Since we've already shown $\lambda_{\kappa}(G_{n,r}) \geq \lambda_{\kappa}(H)$ with probability $1 - o(n^{-1})$, this completes the proof of Lemma~\ref{lem:higher_order_eigenvalue_tail_bound}.
+
+\subsection{Proof of Lemma~\ref{lem:higher_order_eigenvalue_tail_bound}}
+
+Let $D_{\min} = \min_{i \in [n]} D_{ii}$ be the minimum degree of any sample $x \in X$ in the graph $G_{n,r}$. We have that
+\begin{equation*}
+\lambda_{\kappa}(L) \geq D_{\min} - \lambda_{n - \kappa}(A)
+\end{equation*}
+We
+
 \end{document}

### 2020-01-25 notes/weekly_notes/12.20.19.NotesForWeek.tex
updated weekly notes and project level notes

@@
 \newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
-\newcommand{\gradient}{\mathcal{D}}
+\newcommand{\gradient}{\nabla}
 \newcommand{\Hessian}{\mathcal{D}^2}
 \newcommand{\dotp}[2]{\langle #1, #2 \rangle}
+\newcommand{\deriv}{\mathcal{D}}
 
@@
 
-\section{Bounding the Roughness Functional in Expectation}
+\section{Bounding the Roughness Functional when $f$ is Holder}
 
-For an integer $\ell > 0$, we will say $K$ is a $2$nd-order kernel if it is bounded and compactly supported on $B(0,1)$ and additionally
+For an integer $\ell > 0$, we will say $K$ is a $2$nd-order kernel if it is compactly supported on $B(0,1)$, uniformly bounded $\abs{K(x)} \leq K_{\max}$ for all $x \in B(0,1)$, and additionally
 \begin{equation*}
@@
 
+\section{Bounding the Roughness Functional when $f$ is Sobolev}
+Suppose $f$ is assumed to belong to the Sobolev space $W_d^{s,2}(\mathcal{X})$, and have small semi-norm in this space, rather than belonging to the Holder space $C_d^{s}(\Xset)$ and having small Holder norm.
+Can we still prove--possibly up to constants--the same bound on the expected graph semi-norm of $f$?
+Before answering this question, we review the definition of Sobolev spaces and their associated norms.
+
+\subsection{Sobolev spaces}
+
+For a given $s > 0$, the Sobolev space $W_d^{s,2}(\mathcal{X})$ consists of all functions $f \in \mathcal{L}^2(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^2(\mathcal{X})$.
+The Sobolev $\{s,2\}$ semi-norm is then
+\begin{equation*}
+[f]_{W_d^{s,2}(\mathcal{X})} = \left(\sum_{\abs{\alpha} = s} \int_{\Xset} \abs{D^{\alpha}f}^2 \,dx\right)^{1/2}
+\end{equation*}
+and the Sobolev $\{s,2\}$ norm is 
+\begin{equation*}
+\norm{f}_{W_d^{s,2}(\mathcal{X})}^2 = \left(\sum_{q = 1}^{s} [f]_{W_d^{s,2}(\mathcal{X})}^2\right)^{1/2}.
+\end{equation*}
+For a $R > 0$, the corresponding ball is $W_d^{s,2}(\Xset; R) = \set{f: \norm{f}_{W^{s,2}(\Xset)} \leq R}$.
+
+\subsection{Bounds on the graph semi-norm}
+
+Under largely the same conditions as Lemma~\ref{lem:roughness_functional_expectation}, we have that the roughness functional $R_{s,n}(f)$ is (up to constants), no greater than the Sobolev norm $\norm{f}_{W_d^{1,2}(\mathcal{X})}$.
+\begin{lemma}
+	\label{lem:roughness_functional_expectation_sobolev}
+	Let $\Xset$ be a Lipschitz domain.
+Suppose that $f \in W^{s,2}(\Xset)$, and further that $p \in C^{s-1}(\Xset;p_{\max})$ for some constant $p_{\max}$.
+Then for any $2$nd-order kernel $K$ and any $n^{-1/(2(s - 1) + d)} \leq r(n) \leq 1$, for sufficiently large $n$ the expected graph Sobolev seminorm is upper bounded
+	\begin{equation}
+	\label{eqn:roughness_functional_expectation_sobolev}
+	\Ebb\bigl[R_{s,n}(f)\bigr] \leq c \cdot \norm{f}_{W_d^{s,2}(\Xset)}
+	\end{equation}
+	for some constant $c$ which may depend on $s$, $p_{\max}$, $K_{\max}$, $d$ and $\Xset$, but not on $f$, $r$ or $n$.
+\end{lemma}
+
+Note that the bound~\eqref{eqn:roughness_functional_expectation_sobolev} involves, on the right hand side, the norm $\norm{f}_{W_d^{s,2}(\Xset)}$ as opposed to the seminorm $[f]_{W_d^{s,2}(\Xset)}$.
+To better understand this, consider the following operator
+\begin{equation*}
+L_rf(x) = \frac{1}{r^2}\int (f(z) - f(x))K_r(z,x) \,dP(x)
+\end{equation*} 
+The operator $L_r$ is the expectation of the graph Laplacian, in the sense that $\Ebb(L_nf(x)) = L_rf(x)$, and so it makes sense that the behavior of the associated seminorm $\dotp{L_r^sf}{f}$ is related to the behavior of $\dotp{L_n^sf}{f}$.
+However, for any $s$ the seminorm $\dotp{L_r^sf}{f}$ assigns zero length only to constant functions $f$.
+As a one-dimensional example, if $s = 2$, and $f(x) = c \cdot x$ for $c$ large, the seminorm $\dotp{L_r^sf}{f}$ will also be quite large, despite the fact that the second derivative $f'' = 0$.
+
+To be clear, $\Ebb[\dotp{L_n^sf}{f}] \neq \dotp{L_r^sf}{f}$, and bounding the former turns out to be non-trivial.
+The proof of Lemma~\ref{lem:roughness_functional_expectation_sobolev} relies on Lemma~\ref{lem:expected_difference_operators_sobolev}.
+
+\begin{lemma}
+	\label{lem:expected_difference_operators_sobolev}
+	Let $\Xset$ be a Lipschitz domain, and suppose $f \in W_d^{s,2}(\Xset)$ for some $s \in \mathbb{N}_{+}$.
+Let $q = s/2$ when $s$ is even, and $q = (s - 1)/2$ when $s$ is odd.
+For any indices $k = (k_1,\ldots,k_q)$ and $\ell = (\ell_1,\ldots,\ell_q)$, we have
+	\begin{equation}
+	\label{eqn:expected_difference_operators_sobolev_1}
+	\Ebb(D_kf(x_i) D_\ell f(x_i)) =
+	\begin{cases*}
+	O(r^{2s}) \cdot \norm{f}_{W_d^{s,2}(\Xset)}^2, & ~~\textrm{if all indices are distinct} \\
+	O(r^{2} r^{d(\abs{k \cup \ell \cup i} - (2q + 1))})\cdot [f]_{W_d^{1,2}(\Xset)}^2, & ~~\textrm{otherwise}~ 
+	\end{cases*}
+	\end{equation}
+	Additionally, we have
+	\begin{equation}
+	\label{eqn:expected_difference_operators_sobolev_2}
+	\Ebb(d_iD_kf(x_j) d_iD_\ell f(x_j)) =
+	\begin{cases*}
+	O(r^{2s}) \cdot \norm{f}_{W_d^{s,2}(\Xset)}^2, & ~~\textrm{if all indices are distinct} \\
+	O(r^{2} \cdot r^{d(\abs{k \cup \ell \cup i \cup j} - (2q + 2))}) \cdot [f]_{W_d^{1,2}(\Xset)}^2, & ~~\textrm{otherwise}~ 
+	\end{cases*}
+	\end{equation}
+\end{lemma}
+
+The lower bound on $r(n)$ in Lemma~\ref{lem:roughness_functional_expectation_sobolev} is needed to ensure the leading term in~\eqref{eqn:expected_difference_operators_sobolev_1} (or~\eqref{eqn:expected_difference_operators_sobolev_2}), where all indices are distinct, dominates the lower-order terms, where indices are repeated.
+
 \section{Proofs}
+
+\subsection{Proof of Lemma~\ref{lem:expected_difference_operators_sobolev}}
+
+We first prove the desired bound in the case when some indices are repeated, and then the desired bound in the case when all indices are distinct.
+
+\subsubsection{Repeated indices.}
+
+Since the proofs of~\eqref{eqn:expected_difference_operators_sobolev_1} and~\eqref{eqn:expected_difference_operators_sobolev_2} are essentially the same for the case where some index is repeated, we will assume without loss of generality that $s$ is even.
+Let $k,\ell \in [n]^q$ be index vectors for $q = s/2$.
+
+When at least one index is repeated, we obtain a sufficient upper bound by reducing the problem of upper bounding the iterated difference operator to that of upper bounding a single difference operator.
+Letting $k = (k_1,\ldots,k_q)$, we can show by induction that the absolute value of the iterated difference operator $\abs{D_kf(x_i)}$ is upper bounded by
+\begin{equation*}
+\abs{D_kf(x_i)} \leq \left(\frac{2K_{\max}}{r^d}\right)^{q-1} \sum_{h \in k \cup i} \abs{D_{k_q}f(x_h)} \cdot \1\{G_{n,r}[X_{k \cup i}]~\textrm{is a connected graph} \}.
+\end{equation*}
+Therefore,
+\begin{align}
+\abs{D_kf(x_i)} \abs{D_{\ell}f(x_i)} & \leq \left(\frac{2K_{\max}}{r^d}\right)^{2(q - 1)} \sum_{h,j \in k \cup \ell \cup i} \abs{D_{k_q}f(x_h)} \cdot \abs{D_{\ell_q}f(x_j)} \cdot \1\{G_{n,r}[X_{k \cup i}], G_{n,r}[X_{\ell \cup i}]~\textrm{are connected graphs.} \} \nonumber \\
+& =  \left(\frac{2K_{\max}}{r^d}\right)^{2(q - 1)} \sum_{h,j \in k \cup \ell \cup i} \abs{D_{k_q}f(x_h)} \cdot \abs{D_{\ell_q}f(x_j)} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is a connected graph.} \} \label{eqn:expected_difference_operators_sobolev_pf0}
+\end{align}
+
+We now break our analysis into three cases, based on the number of distinct indices in $k_q,\ell_q,h,j$.
+
+\paragraph{Case 1: Two distinct indices.}
+Let $k_q = \ell_q = i$, and $h = j$. Using the law of iterated expectation, we obtain
+\begin{align*}
+\Ebb \left[\left(D_{i}f(x_j)\right)^2 \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\right] &= \Ebb \left[\left(D_{i}f(x_j)\right)^2 \cdot \Pbb\left[\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \} |x_i,x_j\right]\right] \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 2)d}) \cdot \Ebb\left[(D_i(f(x_j))^2\right] \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 3)d}) \cdot \Ebb\left[(d_if(x_j))^2K_r(x_i,x_j)\right] \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) [f]_{W^{1,2}(\Xset)}^2
+\end{align*}
+where the last equality follows from Lemma~\ref{lem:expected_first_order_seminorm}.
+\paragraph{Case 2: Three distinct indices.}
+Let $k_q = \ell_q = i$, for some $i \neq j \neq h$. Using the law of iterated expectation, we obtain
+\begin{align}
+\Ebb & \left[\abs{D_{i}f(x_j)} \cdot \abs{D_if(x_h)} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\right] \\
+& = \Ebb \left[\abs{D_{i}f(x_j)} \cdot \abs{D_if(x_h)} \cdot \Pbb\left[\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \} |x_i,x_j,x_h\right]\right] \nonumber \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 3)d}) \cdot \Ebb\left[\abs{D_{i}f(x_j)} \cdot \abs{D_if(x_h)} \right]. \label{eqn:expected_difference_operators_sobolev_pf1}
+\end{align}
+It remains to upper bound $\Ebb\left[\abs{D_{i}f(x_j)} \cdot \abs{D_if(x_h)} \right]$, which we rewrite as
+\begin{align*}
+\Ebb\left[\abs{D_{i}f(x_j)} \cdot \abs{D_if(x_h)} \right] & = \int \int \int \abs{f(z) - f(x)} \abs{f(z) - f(y)} K_r(z,y) K_r(z,x) \,dP(x) \,dP(y) \,dP(x) \\
+& = \int \left[\int \abs{f(z) - f(x)} K_r(z,x) \,dP(x)\right]^2 \,dP(z) \\
+& \leq p_{\max}^3 \int_{\Xset} \left[\int_{\Xset} \abs{f(z) - f(x)} K_r(z,x) \,dx\right]^2 \,dz
+\end{align*}
+By the argument set out in Section~\ref{subsec:proof_of_lemma_graph_Sobolev_seminorm_1}, we may assume without loss of generality that there exists an extension $g \in C_d^{\infty}(\Xset) \cap W_d^{1,2}(\Rd)$ such that $[g]_{W_d^{1,2}(\Rd)} \leq c [f]_{W_d^{1,2}(\Xset)}$.
+Then, by a similar series of steps as in Section~\ref{subsec:proof_of_lemma_graph_Sobolev_seminorm_1}, we obtain
+\begin{align*}
+\int_{\Xset} \abs{f(z) - f(x)} K_r(z,x) \,dx & \leq \int_{\Rd} \abs{g(z) - g(x)} K_r(z,x) \,dx \\
+& = \int_{\Rd} \abs{\int_{0}^{1} \dotp{\nabla g(x + t(z - x))}{z - x} \,dt} K_r(z,x) \,dx \\
+& \leq \int_{\Rd} \int_{0}^{1} \norm{\nabla g(x + t(z - x))}\cdot\norm{z - x} \,dt K_r(z,x) \,dx \\
+& \leq r \int_{\Rd} \int_{0}^{1} \norm{\nabla g(x + t(z - x))} \,dt K_r(z,x) \,dx \\
+& \leq r \frac{K_{\max}}{r^d} \int_{B(z,r)} \int_{0}^{1} \norm{\nabla g(x + t(z - x))} \,dt  \,dx \\
+& \leq r K_{\max} \int_{B(0,1)} \int_{0}^{1} \norm{\nabla g(x - try)} \,dt  \,dy,
+\end{align*}
+and as a result, 
+\begin{equation*}
+p_{\max}^3 \int_{\Xset} \left[\int_{\Xset} \abs{f(z) - f(x)} K_r(z,x) \,dx\right]^2 \,dz \leq c\cdot p_{\max}^3 r^2 K_{\max}^3 [f]_{W_d^{1,2}(\Rd)}^2 = O(r^2)\cdot [f]_{W_d^{1,2}(\Xset)}^2
+\end{equation*}
+Plugging back into~\eqref{eqn:expected_difference_operators_sobolev_pf1}, we have that $\Ebb \left[\abs{D_{i}f(x_j)} \cdot \abs{D_if(x_h)} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\right] = O(r^{d(\abs{k \cup \ell \cup i} - 3) + 2}) \cdot [f]_{W_d^{1,2}(\Xset)}^2$.
+
+\paragraph{Case 3: Four distinct indices.}
+Using the law of iterated expectation, we find that
+\begin{align*}
+\Ebb[ &\abs{D_{k_q}f(x_i)}{\abs{D_{\ell_q}f(x_j)}} \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}] \\
+& = \Ebb[\abs{D_{k_q}f(x_i)}{\abs{D_{\ell_q}f(x_j)}} \Pbb(G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected}|x_i,x_j,x_{k_q},x_{\ell_q})] \\
+& = O(r^{d(\abs{k \cup \ell \cup i} - 4)}) \cdot \Ebb\bigl[\abs{D_{k_q}f(x_i)}\cdot{\abs{D_{\ell_q}f(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\bigr] 
+\end{align*}
+Replacing $f$ by a smooth extension $g \in C^{\infty}(\Xset) \cap W_d^{1,2}(\Rd)$, we rewrite the expectation as an integral,
+\begin{align*}
+\Ebb\bigl[& \abs{D_{k_q}f(x_i)}\cdot{\abs{D_{\ell_q}f(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\bigr] \\
+& \leq p_{\max}^4 \int_{\Xset^4} \abs{g(x) - g(y)} \abs{g(u) - g(v)} K_r(x,y) K_r(u,v) \1\{\norm{y - v} \leq (2q + 1)r\} \,dy \,dx \,du \,dv
+\end{align*}
+By substituting $z_1 = (y - v)/r$, $z_2 = (u - v)/r$, and $z_3 = (x - y)/r = (x - v)/r + z_1$, we can simplify the integral in the previous display,
+\begin{align*}
+\int_{\Xset^4} & \abs{g(x) - g(y)} \abs{g(u) - g(v)} K_r(x,y) K_r(u,v) \1\{\norm{y - v} \leq (2q + 1)r\} \,dy \,dx \,du \,dv \\
+& \leq K_{\max}^2 r^d \int_{\Xset} \int_{[B(0,1)]^3} \abs{g\bigl((z_3 + z_1)r + v\bigr) - g(z_1r + v)} \bigl|g(z_2r + v) - g(v)\bigr| \,dz_1 \,dz_2 \,dz_3 \,dv \\
+& \leq  K_{\max}^2 r^{d + 2} \int_{[B(0,1)]^3} \int_{[0,1]^2} \int_{\Xset} \norm{\nabla g(t z_3 r + z_1r + v)} \cdot \norm{\nabla g(t z_2 r + v)} \,dv \,dt_1 \,dt_2 \,dz_1 \,dz_2 \,dz_3 \\
+& \leq c \nu_d^3 K_{\max}^2 r^{d + 2} [f]_{W_{d}^{1,2}(\Xset)}^2.
+\end{align*}
+Therefore $\Ebb[ \abs{D_{k_q}f(x_i)}{\abs{D_{\ell_q}f(x_j)}} \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}] = O(r^{d(\abs{k \cup \ell \cup i} - 3) + 2})\cdot [f]_{W_d^{1,2}(\Xset)}^2$.
+
+Since we obtain the same scaling in each case, plugging this back in to~\eqref{eqn:expected_difference_operators_sobolev_pf0} we have that for any $k, \ell \in [n]^q$
+\begin{equation*}
+\Ebb\bigl[\abs{D_kf(x_i)} \abs{D_\ell f(x_i)}\bigr] = O(r^{d(\abs{k \cup \ell \cup i} - (2q + 1)) + 2}) \cdot [f]_{W_d^{1,2}(\Xset)}^2.
+\end{equation*}
+
+\subsubsection{All indices distinct.}
+
+We first show the desired result when $s$ is even, and then when $s$ is odd.
+In either case, since $\Xset$ is a Lipschitz domain, there exists an extension $g \in W^{s,2}(\Rd)$ such that $\norm{g}_{W^{s,2}(\Rd)} \leq c \norm{f}_{W^{s,2}(\Xset)}$, and $g = f$ a.e in $\Xset$.
+Further, there exists a sequence of smooth approximations $(g_m) \subset C^{\infty}(\Rd)$, such that
+\begin{equation*}
+\norm{g_m - g}_{W^{s,2}(\Rd)} \to 0.
+\end{equation*}
+
+\paragraph{$s$ even.}
+
+By the law of iterated expectation, we have
+\begin{align*}
+\Ebb\bigl[D_kf(x_i)D_kf(x_j)\bigr] = \Ebb\bigl[D_kg(x_i)D_kg(x_j)\bigr] & = \Ebb\Bigl[\bigl(\Ebb[D_kg(x_i)|x_i]\bigr)^2\Bigr] \\
+& \leq 2\Ebb\Bigl[\bigl(\Ebb[D_kg_m(x_i)|x_i]\bigr)^2\Bigr] + \Ebb\Bigl[\bigl(\Ebb[D_k(g - g_m)(x_i)|x_i]\bigr)^2\Bigr] \\
+& \leq 2\Ebb\Bigl[\bigl(\Ebb[D_kg_m(x_i)|x_i]\bigr)^2\Bigr] + \left(\frac{2K_{\max}}{r^d}\right)^q \norm{g - g_m}_{\Leb^2(\Rd)}^2.
+\end{align*}
+By Lemma~\ref{lem:leading_term_sobolev}, we have that $\Ebb[D_kg_m(x_i)|x_i]\bigr)^2 = O(r^{2s})\cdot [f_s(x)]^2$ for some $f_s \in \Leb^2(\Rd)$ which satisfies $\norm{f_s}_{\Leb^2(\Rd)} \leq c \norm{g_m}_{W^{s,2}(\Rd)}$ for some constant $c$ which does not depend on $g_m$.
+Along with the previous display, this implies
+\begin{equation*}
+\Ebb\bigl[D_kf(x_i)D_kf(x_j)\bigr] \leq O(r^{2s}) \norm{g_m}_{W^{s,2}(\Rd)}^2 + \left(\frac{2K_{\max}}{r^d}\right)^q \norm{g - g_m}_{\Leb^2(\Rd)}^2,
+\end{equation*}
+and taking the limit of the right hand side as $m \to \infty$ we have $\Ebb\bigl[D_kf(x_i)D_kf(x_j)\bigr] \leq O(r^{2s})\cdot\norm{f}_{W^{s,2}(\Rd)}^2$.
+
+\paragraph{$s$ odd.}
+
+By the law of iterated expectation, we have
+\begin{align*}
+\Ebb[d_iD_kg_m(x_j)d_iD_{\ell}g_m(x_j)K_r(x_i,x_j)] & = \Ebb\biggl[\Bigl(d_i\bigl(\Ebb(D_kf)\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \\
+& = \Ebb\biggl[\Bigl(d_i\bigl(I_{s - 1}\cdot f_{s - 1} + O(r^s)f_s(x_j)\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr].
+\end{align*}
+where the latter equality follows from Lemma~\ref{lem:leading_term_sobolev}.
+Here $I_{s - 1}, f_{s -1}$ and $f_s$ satisfy the conclusions of that Lemma, namely that $\abs{I_{s - 1}} \leq r^{s - 1}$, $f_{s - 1}$ and $f_s \in C^{\infty}(\Rd)$, and
+\begin{equation*}
+\norm{f_{s - 1}}_{W^{1,2}(\Rd)}, \norm{f_s}_{\Leb^2(\Rd)} \leq c \norm{g_m}_{W^{s,2}(\Rd)}.
+\end{equation*}
+By the linearity and boundedness of the difference operator $d_i$, we have
+\begin{align*}
+\Ebb\biggl[\Bigl(d_i\bigl(I_{s - 1}\cdot f_{s - 1} + O(r^s)f_s(x_j)\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] & = \Ebb\biggl[\Bigl(I_{s - 1} d_if_{s - 1}(x_j) + O(r^s)f_s(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \\
+& \leq 2I_{s - 1}^2 \Ebb\Bigl[\bigl(d_if_{s-1}(x_j)\bigr)^2K_r(x_i,x_j)\Bigr] + O(r^{2s})\Ebb\Bigl[f_s(x_j)^2\Bigr] \\
+& = O(r^{2(s - 1)})\Ebb\Bigl[\bigl(d_if_{s-1}(x_j)\bigr)^2K_r(x_i,x_j)\Bigr] + O(r^{2s}) \norm{g_m}_{W^{s,2}(\Rd)}^2 \\
+& \leq O(r^{2s}) \norm{g_m}_{W^{s,2}(\Rd)}^2
+\end{align*}
+where the last inequality follows from Lemma~\ref{lem:expected_first_order_seminorm}.
+By similar arguments to the case where $s$ is even, we can show that this estimate holds (up to constants) when $g_m$ is replaced by $f$, establishing the claim.
 
@@
 establishing $\Ebb(d_iD_kf(x_j)d_iD_{\ell}f(x_j)K_r(x_i,x_j)) = O(r^{2s})$ as claimed.
+
 \section{Auxiliary Results.}
@@
 
-	We will prove by induction on $q$ in the case where $d = 1$. The When $q = 1$ and $k \in [n]$, by taking Taylor expansions of $f$ and $p$, we obtain
+	We will prove by induction on $q$ in the case where $d = 1$. When $q = 1$ and $k \in [n]$, by taking Taylor expansions of $f$ and $p$, we obtain
 	\begin{align}
@@
 \end{proof}
+
+The following Lemma establishes a result analogous to Lemma~\ref{lem:leading_term} when the function $f$ and density $p$ are assumed to be bounded in Sobolev, rather than Holder, norm.
+We use multiindex notation to represent higher order partial derivatives and polynomials.
+For $\alpha \in [\mathbb{N}]^d$, and $x,z \in \Rd$ we write
+\begin{equation*}
+\abs{\alpha} = \alpha_1 + \cdots + \alpha_d,~~ f^{(\alpha)} = \frac{\partial^{\abs{\alpha}}f}{\partial x_{\alpha_1}\ldots \partial x_{\alpha_q}},~~ (x - z)^{\alpha} := (x_{\alpha_1} - z_{\alpha_1})\cdots(x_{\alpha_d} - z_{\alpha_d})
+\end{equation*}
+\begin{lemma}
+	\label{lem:leading_term_sobolev}
+	Let $k \in [n]^q$ for some $q \geq 1$. Suppose that $f \in W_d^{s,2}(\Rd) \cap C_d^{\infty}(\Rd)$ and $p \in C_d^{s - 1}(\Rd;p_{\max})$, and that $K_r$ is a second order kernel. Then there exist
+	\begin{itemize}
+		\item functions $f_{\ell}, \ell = 2q,\ldots,s$ satisfying $f_{\ell} \in W_d^{s - \ell,2}(\Rd) \cap C_d^{\infty}(\Rd)$ and
+		\begin{equation*}
+		\norm{f_{\ell}}_{W_d^{s - \ell,2}(\Rd)} \leq c \norm{f}_{W_d^{s,2}(\Rd)}
+		\end{equation*}
+		for some constant $c$ which depends only on $d$, $\Xset$ and $p_{\max}$, and 
+		\item constants $I_{\ell},\ell = 2q,\ldots,s$ which depend only on $K(\cdot)$ satisfying $\abs{I_{\ell}} \leq r^{\ell}$,
+	\end{itemize}
+	such that
+	\begin{equation}
+	\label{eqn:leading_term_sobolev}
+	\Ebb(D_kf(x)) =
+	\begin{cases*}
+	\sum_{\ell = 2q}^{s - 1} I_{\ell} \cdot f_{\ell}(x) +  O(r^{s}) \cdot f_{s}(x),~~& \textrm{if $2q < s$} \\
+	O(r^{s}) \cdot f_{s}(x),~~& \textrm{if $2q \geq s$}
+	\end{cases*}
+	\end{equation}
+	The $O(\cdot)$ term may depend on $s, K_{\max}$ and $p_{\max}$, but does not depend on $f$.
+\end{lemma}
+\begin{proof}	
+	We proceed by induction on $q$. 
+	\paragraph{Base case.}
+	We begin with the base case of $q = 1$. Since $f$ and $p$ are smooth, they both admit Taylor expansions of the following form for all $x,z \in \Rd$:
+	\begin{align*}
+	f(z) & = \sum_{\abs{\alpha} < s} \frac{f^{(\alpha)}(x)}{\alpha!} (x - z)^{\alpha} + \frac{\abs{\alpha}}{\alpha!}\sum_{\abs{\alpha} = s} (x - z)^{\alpha} \int_{0}^{1}(1 - t)^{s - 1} f^{(\alpha)}(x + t(z - x)) \,dt\\
+	p(z) & = \sum_{\abs{\beta} < s  - 1} \frac{p^{(\beta)}(x)}{\beta!} (x - z)^{\beta} + O((x - z)^{s - 1})
+	\end{align*}
+	where $f^{(\alpha)} \in W^{s - \abs{\alpha},2}(\Rd) \cap C^{\infty}(\Rd)$ additionally satisfies
+	\begin{equation*}
+	\norm{f^{(\alpha)}}_{W^{s - \abs{\alpha},2}(\Rd)} \leq \norm{f}_{W^{s,2}(\Rd)}
+	\end{equation*}
+
+	Replacing $f$ by its Taylor expansion inside the expected first order difference operator $\Ebb(D_kf(x))$ and letting $E_{\alpha,P} := \Ebb\left[(x - x_k)^{\alpha}K_r(x_k,x)\right]$, we have
+	\begin{align}
+	\Ebb(D_kf(x)) & = \sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)}{\alpha!} E_{\alpha,P} + \frac{\abs{\alpha}}{\alpha!}\sum_{\abs{\alpha} = s} \int_{0}^{1}(1 - t)^{s - 1} \Ebb\bigl[f^{(\alpha)}(x + t(x_k - x)) (x_k - x)^{\alpha} K_r(x_k,x)\bigr] \,dt \nonumber \\
+	& =  \sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)}{\alpha!} E_{\alpha,P} + O(r^s)\sum_{\abs{\alpha} = s} \int_{0}^{1}\Ebb\bigl[f^{(\alpha)}(x + t(x_k - x))K_r(x_k,x)\bigr] \,dt \nonumber \\
+	& =  \sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)}{\alpha!} E_{\alpha,P} + O(r^s)\sum_{\abs{\alpha} = s} \int_{0}^{1} \int_{B(x,r)}\frac{f^{(\alpha)}(x + t(z - x))}{r^d} \,dz \,dt \nonumber \\
+	& = \sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)}{\alpha!} E_{\alpha,P} + O(r^s)\int_{0}^{1} \int_{B(0,1)} \sum_{\abs{\alpha} = s} f^{(\alpha)}(x + try) \,dy \,dt \label{eqn:leading_term_sobolev_pf1}
+	\end{align}
+	Turning our attention now to the expectation $E_{\alpha,P}$, by replacing $p$ with its Taylor expansion we obtain
+	\begin{equation*}
+	E_{\alpha,P} = \int_{\Rd} (x - z)^{\alpha} K_r(x,z) p(z) \,dz \sum_{\abs{\beta} = 0}^{s - 2} \frac{p^{(\beta)}(x)}{\beta!} \underbrace{\int_{\Rd} (x - z)^{\alpha + \beta}K_r(x,z)\,dz}_{:=I_{\abs{\alpha} + \abs{\beta}}} + O(r^{s})
+	\end{equation*}
+	where the first sum equals zero when $s = 1$. Plugging this back in to~\eqref{eqn:leading_term_sobolev_pf1} yields
+	\begin{align*}
+	\Ebb(D_kf(x)) & = \sum_{\alpha = 1}^{s - 1} \sum_{\abs{\beta} = 0}^{s-2} \frac{f^{(\alpha)}(x)p^{(\beta)}(x)}{\alpha!\beta!} I_{\abs{\alpha} + \abs{\beta}} + \\ & O(r^s)\cdot\biggl(\underbrace{\sum_{\abs{\alpha} = 1}^{s-1}  f^{(\alpha)}(x) + \int_{0}^{1} \int_{B(0,1)} \sum_{\abs{\alpha} = s} f^{(\alpha)}(x + try) \,dy \,dt}_{: = g_s(x)}\biggr)
+	\end{align*}
+	where $g_s \in \Leb^2(\Rd)$ by Lemma~\ref{lem:remainder_term}.
+We are now in a position to prove the second part of~\eqref{eqn:leading_term_sobolev} when $q = 1$.
+Since $q = 1$, $s \in \{1,2\}$.
+When $s = 1$, the sum in the prior expression is over no terms and is equal to zero.
+Since the integral $I_{1} = 0$, the first term is also zero in the case where $s = 2$.
+For $s \in \{1,2\}$, defining $f_s := g_s$, we have $\Ebb(D_kf(x)) = O(r^s) f_s(x)$ for $f_s \in \Leb_d^2(\Xset;R)$.
+This proves the second part of~\eqref{eqn:leading_term_sobolev} when $q = 1$.
+
+	Otherwise when $s > 2$ and $q = 1$, we must analyze the first term in the prior expression.
+Since $f^{(\alpha)} \in W^{s - \abs{\alpha},2}(\Rd)$ and $p^{(\beta)} \in C^{s - 1 - \abs{\beta}}(\Rd)$, and $\min\{s - \abs{\alpha}, s - 1 - \abs{\beta}\} \geq s - (\abs{\alpha} + \abs{\beta})$, the product $f^{(\alpha)} p^{(\beta)}$ belongs to $W^{s - (\abs{\alpha} + \abs{\beta}),2}(\Rd)$, and moreover
+	\begin{equation*}
+	\norm{f^{(\alpha)} p^{(\beta)}}_{W^{s - (\abs{\alpha} + \abs{\beta}),2}(\Rd)} \leq p_{\max} \norm{f^{(\alpha)}}_{W^{s - (\abs{\alpha} + \abs{\beta}),2}(\Rd)} \leq p_{\max} \norm{f}_{W^{s,2}(\Rd)}.
+	\end{equation*} 
+	Additionally, the integral $I_1 = 0$, and for $\ell > 1$,
+	\begin{equation*}
+	\abs{I_{\ell}} \leq r^{\ell} \int K_r(x,z)\,dz = r^{\ell}
+	\end{equation*}
+	Therefore,
+	\begin{align*}
+	\sum_{\alpha = 1}^{s - 1} \sum_{\abs{\beta} = 0}^{s-2} \frac{f^{(\alpha)}(x)p^{(\beta)}(x)}{\alpha!\beta!} I_{\abs{\alpha} + \abs{\beta}} & = \sum_{\ell = 1}^{s - 1} \sum_{\substack{\abs{\alpha} + \abs{\beta} = \ell, \\ \abs{\alpha} > 0}} \frac{f^{(\alpha)}(x)p^{(\beta)}(x)}{\alpha!\beta!} I_{\ell} + O(r^{s}) \sum_{\abs{\alpha} = 2}^{s - 1} \sum_{\abs{\beta} = {s - \abs{\alpha}}}^{s - 2} \frac{f^{(\alpha)}(x)p^{(\beta)}(x)}{\alpha!\beta!} \\
+	& = \sum_{\ell = 2}^{s - 1} I_{\ell} \left( \underbrace{\sum_{\substack{\abs{\alpha} + \abs{\beta} = \ell, \\ \abs{\alpha} > 0}} \frac{f^{(\alpha)}(x)p^{(\beta)}(x)}{\alpha!\beta!}}_{:= f_{\ell}(x)} \right) + O(r^{s}) \underbrace{\sum_{\abs{\alpha} = 2}^{s} \sum_{\abs{\beta} = {s - \abs{\alpha}}}^{s - 1} \frac{f^{(\alpha)}(x)p^{(\beta)}(x)}{\alpha!\beta!}}_{:=h_{s}(x)}
+	\end{align*} 
+	and setting $f_s = g_s + h_s$ implies the first part of~\eqref{eqn:leading_term_sobolev} when $q = 1$.
+
+	\paragraph{Induction Step.}
+	In this part of the proof, the functions $f_{\ell}$ for $\ell = 2,\ldots,s$ and the constants $I_{\ell}$ for $\ell = 2,\ldots,s$ may change from line to line, but will always satisfy the conditions in the theorem statement.
+
+	Now, we assume \eqref{eqn:leading_term_sobolev} holds for all $k \in [n]^q$, and prove the desired estimate on $\Ebb(D_{kj}f(x))$ for each $j \in [n]$.
+By the law of iterated expectation and our analysis of the base case,
+	\begin{align*}
+	\Ebb(D_{kj}f(x)) & = \Ebb(D_k(\Ebb(D_jf))(x)) \\
+	& = \Ebb\left(D_k\left(\sum_{t = 2}^{s - 1} I_t f_t + O(r^s) f_s\right)(x)\right) \\
+	& = \sum_{t = 2}^{s - 1} I_t \cdot \Ebb(D_kf_t(x)) + O(r^s) f_s(x)
+	\end{align*}
+	where the second equality follows from the linearity and boundedness of $f \mapsto \Ebb(D_kf)$.
+We now apply the inductive hypothesis to $\Ebb(D_kf_t(x))$ for each $t = 2,\ldots,s-1$, to prove each part of~\eqref{eqn:leading_term_sobolev}.
+
+	First we consider the case when $2(q + 1) \geq s$.
+Note that $f_t \in W^{s - t,2}(\Rd) \cap C^{\infty}(\Rd)$, and $t \geq 2$ implies $2q \geq s - t$.
+Therefore by hypothesis $\Ebb(D_kf_t(x)) = O(r^{s - t})f_s$ for some $f_s \in \Leb^2(\Rd) \cap C^{\infty}(\Rd)$.
+As a result
+	\begin{equation*}
+	\sum_{t = 2}^{s - 1} I_t \cdot \Ebb(D_kf_t(x)) = \sum_{t = 2}^{s - 1} I_t \cdot O(r^{s - t}) f_s(x) = O(r^s) f_s(x)
+	\end{equation*}
+	establishing that the second part of~\eqref{eqn:leading_term_sobolev} holds for all $q$.
+
+	Otherwise $2(q + 1) < s - 1$.
+For each $t = 2,\ldots, s - 1$, if additionally  $2q \leq s - t - 1$, then by hypothesis $\Ebb(D_kf_t(x)) = \sum_{\ell = 2q}^{s - t - 1} I_{\ell} \cdot f_{\ell + t}(x) + O(r^{s - t}) f_s$, and otherwise $\Ebb(D_kf_t(x)) = O(r^{s - t}) f_s(x)$.
+Therefore,
+	\begin{align*}
+	\sum_{t = 2}^{s - 1} I_t \cdot \Ebb(D_kf_t(x)) & = \sum_{t = 2}^{s - 1 - 2q} I_{t} \cdot \left\{\sum_{\ell = 2q}^{s - t - 1} I_{\ell} \cdot f_{\ell + t}(x) + O(r^{s - t}) \cdot f_s(x) \right\} + \sum_{t = s - 1 - 2q}^{s - 1}I_{t} \cdot O(r^{s - t}) \cdot f_s(x) \\
+	& = \sum_{t = 2}^{s - 1 - 2q} I_t \cdot \left\{\sum_{\ell = 2q}^{s - t - 1} I_{\ell} \cdot f_{\ell + t}(x)\right\}  + O(r^s)\cdot f_s(x) \\
+	& = \sum_{\ell = 2q}^{s - 3} \sum_{t = 2}^{s - \ell - 1} I_{\ell + t} \cdot f_{\ell + t}(x) + O(r^s)\cdot f_s(x).
+	\end{align*}
+	Rewriting the final equation as a sum over $\ell + t = 2(q + 1),\ldots, s - 1$ establishes~\eqref{eqn:leading_term}.
+\end{proof}
+
+\begin{lemma}
+	\label{lem:remainder_term}
+	Suppose $f \in \Leb^2(\Rd)$. Then, the function $g(x) = \int_{0}^{1} \int_{B(0,1)} f(x + aty) \,dy \,dt$ also belongs to $\Leb^2(\Rd)$, with norm
+	\begin{equation*}
+	\norm{g}_{\Leb^2(\Rd)} \leq \nu_d\cdot \norm{f}_{\Leb^2(\Rd)}
+	\end{equation*}
+\end{lemma}
+\begin{proof}
+	We compute the squared norm of $g$,
+	\begin{align*}
+	\norm{g}_{\Leb^2(\Rd)}^2 & = \int_{\Rd} \left(\int_{0}^{1} \int_{B(0,1)} f(x + aty) \,dt \,dy \right)^2 \,dx \\
+	& \leq \nu_d^2 \int_{\Rd} \int_{0}^{1} \frac{1}{\nu_d}\int_{B(0,1)} f^2(x + aty) \,dt \,dy \,dx \tag{Jensen's inequality} \\
+	& = \nu_d^2  \int_{0}^{1} \int_{B(0,1)} \frac{1}{\nu_d}\int_{\Rd}f^2(x + aty) \,dt \,dy \,dx \tag{Fubini's theorem} \\
+	& = \nu_d^2 \norm{f}_{\Leb^2(\Rd)}^2.
+	\end{align*}
+\end{proof}
+
+\begin{lemma}
+	\label{lem:expected_first_order_seminorm}
+	Suppose $g \in C^{\infty}(\Rd)$ and $\abs{p(x)} \leq p_{\max}$ for all $x \in \Rd$. Then
+	\begin{equation*}
+	\Ebb[(g(x_j) - g(x_i))^2K_r(x_i,x_j)] \leq c K_{\max} p_{\max}^2 r^2 [g]_{W^{1,2}(\Rd)}^2
+	\end{equation*}
+	for a constant $c$ which depends only on $\Xset$ and $d$.
+\end{lemma}
+\begin{proof}
+	By the fundamental theorem of calculus we have for any $y,x \in \Rd$,
+	\begin{equation*}
+	g(y) - g(x) = \int_{0}^{1} \frac{d}{dt}\bigl[g(x + t(y - x))\bigr] \,dt = \int_{0}^{1} \dotp{\nabla(g(x + t(y - x)))}{y - x} \,dt
+	\end{equation*}
+	Plugging this into~\eqref{eqn:graph_Sobolev_seminorm_1_pf2}, we obtain
+	\begin{align*}
+	\Ebb[(g(x_j) - g(x_i))^2K_r(x_i,x_j)] & \leq p_{\max^2} \int_{\Rd} \int_{\Rd} (g(y) - g(x))^2 K_r(y,x) \,dy \,dx\\
+	& = p_{\max^2} \int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \dotp{\nabla(g(x + t(y - x)))}{y - x} \,dt\right)^2 K_r(y,x) \,dy \,dx \\
+	& \overset{(i)}{\leq} p_{\max^2} \int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \norm{\nabla(g(x + t(y - x)))}\norm{y - x} \,dt\right)^2 K_r(y,x) \,dy \,dx \\
+	& \overset{(ii)}{\leq} p_{\max^2} r^2\int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \norm{\nabla(g(x + t(y - x)))}\,dt\right)^2 K_r(y,x) \,dy \,dx \\
+	& \overset{(iii)}{\leq} p_{\max^2} r^2\int_{\Rd} \int_{\Rd} \int_{0}^{1} \norm{\nabla(g(x + t(y - x)))}^2 \,dt K_r(y,x) \,dy \,dx \\
+	& \overset{(iv)}{\leq} p_{\max^2} K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(x,r)} \norm{\nabla(g(x + t(y - x)))}^2 \,dy \,dt \,dx \\
+	& \overset{(v)}{\leq}  p_{\max^2} K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(0,r)} \norm{\nabla(g(x + z))}^2  \,dz \,dt \,dx
+	\end{align*}
+	where $(i)$ follows by Cauchy-Schwarz, $(ii)$ follows since either $\norm{y - x} \leq r$ or $K_r(y,x) = 0$, $(iii)$ follows by Jensen's, $(iv)$ follows by the assumption $K \leq K_{\max}$ supported on $B(0,1)$, and $(v)$ follows from the change of variables $z = x + t(y - x)$.
+Finally, again using Fubini's Theorem, we have
+	\begin{align*}
+	K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(0,r)} \norm{\nabla(g(x + z))}^2  \,dz \,dt \,dx & = r^{2 - d}\int_{B(0,r)} \int_{0}^{1} \int_{\Rd} \norm{\nabla(g(x + z))}^2  \,dz \,dt \,dx \\
+	& = K_{\max} r^2 [g]_{W_d^{1,2}(\Rd)}.
+	\end{align*}
+\end{proof}
+
+\clearpage
 
@@
 
+\subsection{First-order graph Sobolev seminorm}
+
+We begin with a bound on the first-order graph semi-norm, under the assumption $f$ has small semi-norm in the first order Sobolev space $W_d^{1,2}(\mathcal{X})$.
+
+\begin{lemma}
+	\label{lem:graph_Sobolev_seminorm_1}
+	Let $b \geq 1$ be a fixed number.
+Suppose $K$ is a uniformly bounded kernel function, compactly supported on $B(0,1)$, $f \in W_d^{1,2}(\Xset)$, and additionally that $p(x) \leq p_{\max}$ for all $x \in \Xset$.
+Then there exists a constant $c$ which may depend only on $p$ such that
+	\begin{equation}
+	\label{eqn:graph_Sobolev_seminorm_1}
+	R_{s,n}(f) \leq c \cdot b \cdot [f]_{W_d^{s,2}(\mathcal{X})}^2 
+	\end{equation}
+	with probability at least $1 - \frac{1}{b}$. 
+\end{lemma}
+
+\subsection{Proof of Lemma~\ref{lem:graph_Sobolev_seminorm_1}}
+\label{subsec:proof_of_lemma_graph_Sobolev_seminorm_1}
+We will bound $\Ebb(R_{s,n}(f)) \leq  c \cdot [f]_{W_d^{s,2}(\mathcal{X})}^2$, whence~\eqref{eqn:graph_Sobolev_seminorm_1} holds with probability at least $1 - \frac{1}{b}$ by Markov's inequality.
+We rewrite
+\begin{equation}
+\label{eqn:graph_Sobolev_seminorm_1_pf1}
+\Ebb(R_{s,n}(f)) = \frac{1}{r^2}\int_{\Xset} \int_{\Xset} (f(y) - f(x))^2 K_r(y,x) p(y) p(x) dy dx \leq \frac{p_{\max}^2}{r^2} \int_{\Xset} \int_{\Xset} (f(y) - f(x))^2 K_r(y,x) \,dy \,dx
+\end{equation}
+
+Since $\Xset \subset \Rd$ is a Lipschitz domain, there exists an extension operator $E: W_d^{1,2}(\Xset) \to W_d^{1,2}(\Rd)$ such that $Ef = f$ a.e.
+on $\Xset$, and $[Ef]_{W_d^{s,2}(\Rd)} \leq c [f]_{W_d^{s,2}(\Xset)}$ for some $c$ which depends only on $\Xset$ and $d$.
+For notational simplicity, we denote this extension $Ef =: g$, and observe that
+\begin{equation}
+\label{eqn:graph_Sobolev_seminorm_1_pf2}
+\int_{\Xset} \int_{\Xset} (f(y) - f(x))^2 K_r(y,x) \,dy \,dx \leq \int_{\Rd} \int_{\Rd} (g(y) - g(x))^2 K_r(y,x) \,dy \,dx.
+\end{equation}
+
+Assume for the moment that $g \in C_d^{\infty}(\Xset_{2r})$, where $\Xset_{2r} = \set{x \in \Rd: \mathrm{dist}(x,\Xset) \leq 2r}$.
+If $g$ is not smooth, we can approximate it by smooth functions $(g_m)$ in such a way that the double integral will not be substantially changed, a fact that we will show at the end of the proof.
+
+If $g \in C_d^{\infty}(\Xset)$, by the fundamental theorem of calculus we have for any $y,x \in \Rd$,
+\begin{equation*}
+g(y) - g(x) = \int_{0}^{1} \frac{d}{dt}\bigl[g(x + t(y - x))\bigr] \,dt = \int_{0}^{1} \dotp{\nabla(g(x + t(y - x)))}{y - x} \,dt
+\end{equation*}
+Plugging this into~\eqref{eqn:graph_Sobolev_seminorm_1_pf2}, we obtain
+\begin{align*}
+\int_{\Rd} \int_{\Rd} (g(y) - g(x))^2 K_r(y,x) \,dy \,dx & = \int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \dotp{\nabla(g(x + t(y - x)))}{y - x} \,dt\right)^2 K_r(y,x) \,dy \,dx \\
+& \overset{(i)}{\leq} \int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \norm{\nabla(g(x + t(y - x)))}\norm{y - x} \,dt\right)^2 K_r(y,x) \,dy \,dx \\
+& \overset{(ii)}{\leq} r^2\int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \norm{\nabla(g(x + t(y - x)))}\,dt\right)^2 K_r(y,x) \,dy \,dx \\
+& \overset{(iii)}{\leq} r^2\int_{\Rd} \int_{\Rd} \int_{0}^{1} \norm{\nabla(g(x + t(y - x)))}^2 \,dt K_r(y,x) \,dy \,dx \\
+& \overset{(iv)}{\leq} K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(x,r)} \norm{\nabla(g(x + t(y - x)))}^2 \,dy \,dt \,dx \\
+& \overset{(v)}{\leq} K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(0,r)} \norm{\nabla(g(x + z))}^2  \,dz \,dt \,dx
+\end{align*}
+where $(i)$ follows by Cauchy-Schwarz, $(ii)$ follows since either $\norm{y - x} \leq r$ or $K_r(y,x) = 0$, $(iii)$ follows by Jensen's, $(iv)$ follows by the assumption $K \leq K_{\max}$ supported on $B(0,1)$, and $(v)$ follows from the change of variables $z = x + t(y - x)$.
+Finally, again using Fubini's Theorem, we have
+\begin{align*}
+K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(0,r)} \norm{\nabla(g(x + z))}^2  \,dz \,dt \,dx & = r^{2 - d}\int_{B(0,r)} \int_{0}^{1} \int_{\Rd} \norm{\nabla(g(x + z))}^2  \,dz \,dt \,dx \\
+& = K_{\max} r^2 [g]_{W_d^{1,2}(\Rd)} \\
+& \leq c K_{\max} r^2 [f]_{W_d^{1,2}(\Xset)},
+\end{align*}
+and plugging back in to~\eqref{eqn:graph_Sobolev_seminorm_1_pf1}, we have $\Ebb(R_{s,n}(f)) \leq  c \cdot [f]_{W_d^{s,2}(\mathcal{X})}^2$.
+
+To complete the proof, it remains to treat the case where $g \not\in C_d^{\infty}(\Xset_{2r})$.
+Since $g \in W_d^{1,2}(\Xset_{2r})$, by Theorem 2 of \textcolor{red}{(Evans)}, there exists a sequence of functions $(g_m)$ such that $g_m \in C_d^{\infty}(\Xset_{2r}) \cap W_d^{1,2}(\Xset_{2r})$ and $\norm{g_m - g}_{W_d^{1,2}(\Xset_{2r})} \to 0$.
+By the triangle inequality,
+\begin{align*}
+\int_{\Xset} \int_{\Xset} (g(y) - g(x))^2 K_r(y,x) \,dy \,dx & \leq 6 \int_{\Xset} \int_{\Xset} (g_m(x) - g(x))^2 K_r(y,x) \,dx + 3 \int_{\Xset} \int_{\Xset} (g_m(y) - g_m(x))^2 K_r(y,x) \,dy \,dx.
+\end{align*}
+The first term shrinks to zero as $m \to \infty$ since
+\begin{equation*}
+\int_{\Xset} \int_{\Xset} (g_m(x) - g(x))^2 K_r(y,x) \,dy \,dx = \int_{\Xset} (g_m(x) - g(x))^2 \,dx \leq \norm{g_m - g}_{W_d^{1,2}(\Xset_{2r})},
+\end{equation*}
+and by our previous arguments, 
+\begin{equation*}
+\limsup_{m \to \infty} \int_{\Xset} \int_{\Xset} (g_m(y) - g_m(x))^2 K_r(y,x) \,dy \,dx \leq \limsup_{m \to \infty} K_{\max}r^2[g_m]_{W_d^{1,2}(\Xset_{2r})} = K_{\max}r^2[g]_{W_d^{1,2}(\Xset_{2r})} \leq c K_{\max}r^2[g]_{W_d^{1,2}(\Xset)}.
+\end{equation*}
+This concludes the proof of Lemma~\ref{lem:graph_Sobolev_seminorm_1}.
+
 \end{document}

### 2020-01-30 notes/graph_testing.tex
added higher-order testing rate overly compactly supported Sobolev functions

@@
 
-We establish some notation that we will use throughout. Let $X = \{x_1,
-\ldots, x_n\}$ be a sample drawn i.i.d. from a distribution $P$ on $\Rd$,
-with density~$p$.  For a radius $r > 0$, we define $G_{n,r}=(V,E)$ to be the
-\emph{$r$-neighborhood graph} of $X$, an unweighted, undirected graph with
-vertices $V=X$, and an edge $(x_i,x_j) \in E$ if and only if $K_r(x_i,x_j) = \norm{x_i -x_j} \leq r$, where $\norm{\cdot}$ is the Euclidean norm. We denote by $A \in
-\Reals^{n \times n}$ the adjacency matrix, with entries $A_{uv} = 1$ if
-$(u,v) \in E$ and $0$ otherwise.  We also denote by $D$ the diagonal degree
-matrix, with entries $D_{uu} := \sum_{v \in V} A_{uv}$. The graph Laplacian is $L = D - A$, and we write its spectral decomposition as $L = V S V^T$.
-
-\section{Regression goodness-of-fit testing with random design.}
-
-Let $P$ be a distribution with density $p$ supported on $\mathcal{X} \subseteq \Reals^d$. Suppose we observe random design points $X = \set{x_1,\ldots,x_n} \sim P$, and additionally responses
+We establish some notation that we will use throughout.
+A graph $G = (V,W)$ consists of a collection of vertices $V = \{v_1,\ldots,v_n\}$, and a weight matrix $W \in \Reals^{\abs{V} \times \abs{V}}$ with $ij$th entry $W_{ij}$ encoding affinity between $v_i$ and $v_j$.
+We will sometimes write $W_G$ to make it clear which graph a weight matrix is associated with.
+When the weight matrix $W_G$ consists only of $0$s and $1$s, we will use the equivalent representation $G = (V,E(G))$ when convenient; here $E(G) = V \times V \cap \{(v_i,v_j): W_{ij} = 1\}$.
+
+We will focus our attention on neighborhood graphs, and operators associated with them.
+For a kernel $K:\Reals \to \Reals$ and radius $r > 0$, we define $G_{n,K}=(V,W)$ to be the neighborhood graph with vertices $V = X$, and weights $W_{ij} = K(\norm{x_i - x_j}/r)$.
+Of particular interest to us will be the \emph{random geometric graph} $G_{n,r}$, the neighborhood graph constructed using the uniform kernel $K(z) := \1\{z \leq 1\}$.
+The weight matrix $W(G_{n,r})$ consists only of $0$s and $1$s, and we will thus use the equivalent representation $G_{n,r} = (X,E(G_{n,r}))$.
+
+The primary operator we are interested in (at this point) is the graph Laplacian. Let $D_G$ be the diagonal degree
+matrix of a graph $G$, with entries $D_{uu} := \sum_{v \in V} W_{uv}$.
+The graph Laplacian is then $L_G = D_G - W_G$.
+Let $\Lambda(G)$ consist of the eigenvalues $0 = \lambda_1(G) \leq \lambda_2(G) \leq \cdots \leq \lambda_n(G)$ of $L_G$, with $v_k(G) = (v_{k,1}(G),\ldots,v_{k,n}(G)) \in \Reals^n$ denoting the eigenvector corresponding to the $k$th eigenvalue $\lambda_k(G)$.
+
+\section{Regression goodness-of-fit testing.}
+
+Suppose we observe samples $(y_i,x_i)$ for $i = 1,\ldots,n$, where conditional on $X$ the responses $Y = \{y_1,\ldots,y_n\}$ are assumed to follow the model
 \begin{equation}
@@
 \end{equation} 
-
-We wish to distinguish
+and our task is to distinguish
 \begin{equation*}
@@
 \end{equation*}
+
 We will evaluate our performance using worst-case risk: for a given function class $\mathcal{H}$ and test function $\phi: \Reals^n \to \set{0,1}$, let
@@
 \paragraph{Eigenvector projection test statistic.}
-Let $VSV^T$ be the spectral decomposition of the Laplacian matrix $L$ of the neighborhood graph $G_{n,r}$.
-To test whether $f = f_0$, we propose the following \emph{eigenvector projection} test statistic:
-\begin{equation}
-\label{eqn:graph_spectral_projections}
-T_{\mathrm{spec}} := \frac{1}{n} \sum_{k = 0}^{\kappa} \left(\sum_{i = 1}^{n} v_{k,i} y_i\right)^2
-\end{equation}
-
-\subsection{Current Results.}
-
-In Theorem~\ref{thm:sobolev_testing_rate} we show that under some typical regularity conditions on $P$, the graph Laplacian eigenvector projection test $\phi_{\textrm{spec}} := \1\{T_{\mathrm{spec}} \geq \tau\}$ is a minimax optimal test over the Sobolev balls $W^{s,2}(\mathcal{X};R)$ whenever $4s > d$ and $\Xset = [0,1]^d$.
+The graph Laplacian eigenvector projection test is a truncated-series test.
+Suppose we are given a graph $G = (X,W)$ defined on the sample points $X$.
+Letting $\kappa$ be some integer between $1$ and $n$, our test projects $Y$ onto the eigenvectors $v_1(G),\ldots,v_{\kappa}(G)$, and then takes the empirical norm of this projection to be the test statistic.
+Formally,
+\begin{equation*}
+T_{\mathrm{spec}}(G) := \frac{1}{n} \sum_{k = 1}^{\kappa} \Biggl(\sum_{i = 1}^{n} y_i v_{k,i}(G)\Biggr)^2
+\end{equation*}
+Then the test $\phi_{\spec}(G) := \1\{T_{\spec}(G) \geq \tau\}$ rejects the null hypothesis when the test statistic $T_{\mathrm{spec}}(G)$ is greater than a pre-specified cutoff $\tau$.
+
+\subsection{Random design.}
+In the random design case, we assume that each design point $x_i$ is sampled independently from some distribution $P$. Formally, our model for the samples is the following:
+\begin{equation}
+\label{eqn:regression_random_design_known_variance}
+x_1,\ldots,x_n \overset{\textrm{i.i.d}}{\sim} P,~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)~\textrm{and}~ \varepsilon \perp X,~~ y_i = f(x_i) + \varepsilon_i.
+\end{equation}
+
+In Theorem~\ref{thm:sobolev_testing_rate_order1} we show that under some typical regularity conditions on $P$, the neighborhood graph Laplacian eigenvector projection test $\phi_{\textrm{spec}}(G_{n,r})$ is a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};R)$ for $d = 1,2,3$ and $\Xset = [0,1]^d$.
+Let $p_d$ equal $3/4$ when $d = 2$ and equal $1/d$ for $d \geq 3$.
+
+\begin{theorem}
+	\label{thm:sobolev_testing_rate_order1}
+	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}.
+Let $R > 0$ and $b \geq 1$ be fixed constants, and $d = 1,2$ or $3$.
+Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p(x)$ is bounded away from zero and infinity,
+	\begin{equation*}
+	0 < p_{\min} < p(x) < p_{\max} < \infty,~~ \textrm{for all $x \in \mathcal{X}$.}
+	\end{equation*}
+	and the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choices
+	\begin{equation*}
+	c \frac{(\log n)^{p_d}}{n^{1/d}} \leq r(n) \leq n^{-4/((4 + d)(2+d))}, ~\kappa = (nR)^{2d/(4 + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	\end{equation*}
+	for $c$ a constant that depends only on $\Xset, p_{\min}$ and $p_{\max}$.
+Then the following statements holds for every $n$ sufficiently large: there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
+	\begin{equation}
+	\label{eqn:sobolev_testing_rate_order1}
+	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4 + d)}\} \cdot n^{-4/(4 + d)}
+	\end{equation}
+	the worst-case risk is upper bounded
+	\begin{equation}
+	\label{eqn:sobolev_testing_rate_order1_1}
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}(G_{n,r}); W^{1,2}(\mathcal{X};R)) \leq \frac{c_2}{b}.
+	\end{equation}
+\end{theorem}
+
+The testing problem in the random design case with known variance fundamentally changes in the low-smoothness regime, where the number of bounded derivatives is small compared to the dimension $d$.
+In the case where $f \in W^{1,2}(\Xset;R)$ is assumed to have one derivative which is bounded in $\Leb^2(\Xset)$ norm, this transition occurs at $d = 4$.
+We shall discuss testing in the low-smoothness regime shortly.
+
+Theorem~\ref{thm:sobolev_testing_rate_order1} establishes that a truncated series test using graph eigenvectors performs just as well, in a minimax sense, as a truncated series test using a classic Fourier series; for analysis of the latter test as well as derivation of the minimax rate of nonparametric goodness-of-fit testing over Sobolev classes see \textcolor{red}{(Ingster)}.
+Intuitively, we can understand this by thinking of the graph eigenvectors as estimates of eigenfunctions of a density-weighted continuum Laplacian operator.
+When the density is uniform, and the domain $\Xset$ is the unit cube, under appropriate boundary conditions these eigenfunctions are themselves elements of the Fourier basis.
+When the density is merely close to uniform, the eigenfunctions no longer belong to the Fourier basis; nevertheless, they share enough properties with the latter that the resulting minimax rate is unchanged.
+Theorem~\ref{thm:sobolev_testing_rate_order1} shows that the additional error due to using \textit{estimates} of these eigenfunctions does not fundamentally change the minimax rate.
+
+When we additionally believe $W^{s,2}(\Xset)$, we know (again from Ingster) that the minimax testing rate is $n^{-4s/(4s + d)}$.
+Now, however, our estimates of the eigenvectors no longer track the derivatives finely enough at the boundaries of $\Xset$ to establish the same minimax rate over all $W^{s,2}(\mathcal{X})$.
+However, if we assume $f$ and its derivatives are supported only on the interior of $\mathcal{X}$, we recover the usual rate.
+Theorem~\ref{thm:sobolev_testing_rate} presents our formal result, that $\phi_{\textrm{spec}}(G_{n,r})$ is a minimax optimal test over the Sobolev balls $W_{\sigma}^{s,2}(\mathcal{X};R)$ whenever $4s > d$ and $\Xset = [0,1]^d$.
 
@@
 	\label{thm:sobolev_testing_rate}
-	Let $b \geq 1$ and $R > 0$ be fixed constants.
-Suppose that $\Pbb$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p \in C^{s-1}(\Xset;p_{\max})$ for some $p_{\max} < \infty$, and further that $p(x)$ is bounded away from zero, i.e.
+	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}.
+Let $R,\sigma > 0$, $b,s,d \geq 1$ be fixed constants, with $s$ and $d$ integers.
+Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p \in C^{s-1}(\Xset;p_{\max})$ for some $p_{\max} < \infty$, and further that $p(x)$ is bounded away from zero, i.e.
 there exists $p_{\min} > 0$ such that
@@
 	\end{equation*}
-	Then the following statement holds: if the test $\phi_{\spec}$ is performed with parameter choices 
-	\begin{equation*}
-	\textrm{$K$ the uniform kernel,}~ n^{-1/(2(s-1) + d)} \leq r(n) \leq n^{-4/((4s + d)(2+d))}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	Then the following statement holds for all $n$ sufficiently large: if the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choice
+	\begin{equation*}
+	n^{-1/(2(s-1) + d)} \leq r(n) \leq \min\bigl\{n^{-4/((4s + d)(2+d))},\sigma/s\bigr\}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
 	\end{equation*}
@@
 	\label{eqn:sobolev_testing_rate_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W^{2,s}(\mathcal{X};R)) \leq \frac{c_2}{b}.
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W_{\sigma}^{s,2}(\mathcal{X};R)) \leq \frac{c_2}{b}.
 	\end{equation}
@@
 Note that when $4s > d$ the critical radius~\eqref{eqn:sobolev_testing_rate} is smaller than \eqref{eqn:L4_testing_rate}.
+
+\subsection{Analysis}
+
+To prove Theorem~\ref{thm:sobolev_testing_rate}, we show that there exists a high-probability set $E \subseteq \Xset^n$ such that conditional on $X \in E$, the test $\phi_{\spec}(G_{n,r})$ has small risk.
+Since $G_{n,r}$ is a function only of $X$ and not of $Y$, this amounts to reasoning about the behavior of the test $\phi_{\spec}$ over a fixed graph $G = (X,E)$, where we observe
+\begin{equation}
+y_i = \beta_i + \varepsilon_i,~~\varepsilon_i \sim \mathcal{N}(0,1)
+\end{equation}
+for some fixed $\beta \in \Reals^n$.
+In Lemma~\ref{lem:fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}(G)$.
+Our bound on the Type II error will be stated as a function of $\beta^T L^s \beta$--a measure of the smoothness the signal $\beta$ displays over the graph $G$--as well as the $\kappa$th eigenvalue $\lambda_{\kappa}$.
+
+\begin{lemma}
+	\label{lem:fixed_graph_testing}
+	Let $1 \leq \kappa \leq n$ be an integer.
+	\begin{enumerate}
+		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}$ is upper bounded
+		\begin{equation}
+		\label{eqn:graph_spectral_type_I_error}
+		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{1}{b^2}.
+		\end{equation}
+		\item \textbf{Type II error:} For any $b \geq 1$ and $\beta$ such that
+		\begin{equation}
+		\label{eqn:fixed_graph_testing_critical_radius}
+		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{\beta^T L^s \beta}{n\lambda_{\kappa}^s}
+		\end{equation}
+		the Type II error of $\phi_{\spec}$ is upper bounded,
+		\begin{equation}
+		\label{eqn:graph_spectral_type_II_error}
+		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{3}{b}.
+		\end{equation}
+	\end{enumerate}
+\end{lemma}
 
@@
 
-For a given $s > 0$, the Sobolev space $W_d^{s,2}(\mathcal{X})$ consists of all functions $f \in \mathcal{L}^2(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^2(\mathcal{X})$.
+For a given $s > 0$, the Sobolev space $W^{s,2}(\mathcal{X})$ consists of all functions $f \in \Leb^2(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^2(\mathcal{X})$.
 The Sobolev $\{s,2\}$ norm is then
 \begin{equation*}
-\norm{f}_{W_d^{s,2}(\mathcal{X})}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^2 \,dx
-\end{equation*}
-and for a given $L > 0$, the corresponding ball is $W_d^{s,2}(\Xset; L) = \set{f: \norm{f}_{W^{s,2}(\Xset)} \leq L}$.
+\norm{f}_{W^{s,2}(\mathcal{X})}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^2 \,dx
+\end{equation*}
+and for a given $R > 0$, the corresponding ball is $W^{s,2}(\Xset; R) = \set{f: \norm{f}_{W^{s,2}(\Xset)} \leq R}$.
+
+For a given $\sigma > 0$, let $\Xset_{\sigma} := \set{x \in \Xset: \mathrm{dist}(x,\partial\Xset) > \sigma}$.
+Let $W_{\sigma}^{s,2}(\Xset;L)$ consist of all those functions $f \in W^{s,2}(\Xset)$ such that $\mathrm{supp}(f) \subseteq \Xset_{\sigma}$.
 
@@
 
-\subsection{Proof of Theorem~\ref{thm:sobolev_testing_rate}}
-
-Our analysis will proceed according to the following steps.
-\begin{enumerate}
-	\item We upper bound the testing error of our eigenvector projection test statistic when $Y$ are viewed as random responses defined over the vertices of a fixed graph $G$.
-This upper bound will hold whenever certain functionals on the graph $G$ are themselves bounded.
-	\item We analyze the behavior of these functionals with respect to the random graph $G_{n,r}$, and bound them with high probability.
-	\item We condition on $X \subseteq \mathcal{E}$, where $\mathcal{E}$ is a high-probability set over which the relevant functionals on $G_{n,r}$ are bounded.
-We conclude that the upper bound on testing error derived in our first step holds whenever $X \subseteq \mathcal{E}$.
-\end{enumerate}
-
-\subsubsection{Step 1: Testing error on a fixed graph}
-
-Let $G = (V,E)$ be a graph over $n$ vertices $V$, and let $\beta = (\beta_1,\ldots,\beta_n) \in \Reals^n$ be a signal over the vertices.
-(Although one can think of $G$ as a realization of $G_{n,r}$ for a particular sample $X$, our theory in this section will deal with $G$ as if it were an arbitrary fixed graph.) We observe responses $Y = (y_1,\ldots,y_n)$ according to the model
-\begin{equation*}
-y_i = \beta_i + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} N(0,1)
-\end{equation*}
-Let $L$ denote the graph Laplacian of $G$, and let $0 = \lambda_1 \leq \lambda_2 \leq \ldots \leq \lambda_n$ be the ordered eigenvalues of $L$, with eigenvector $v_i$ corresponding to eigenvalue $\lambda_i$.
-Our graph spectral test statistic is
-\begin{equation*}
-T_{\spec} = \frac{1}{n}\sum_{k = 1}^{\kappa} \left(\sum_{i = 1}^{n} v_i y_i\right)^2
-\end{equation*}
-where $\kappa$ is a tuning parameter. The resulting test we will use is
-\begin{equation*}
-\phi_{\spec} = \1\{T_{\spec} \geq \frac{\kappa}{n} + t(b)\},~~\textrm{where}~ t(b) = b\sqrt{\frac{2\kappa}{n^2}}~~\textrm{for $b \geq 1$.}
-\end{equation*}
-
-In Lemma~\ref{lem:fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}$.
-Our bound on the Type II error will be stated as a function of $\beta^T L^s \beta$--a measure of the smoothness the signal $\beta$ displays over the graph $G$--as well as the $\kappa$th eigenvalue $\lambda_{\kappa}$.
-
-\begin{lemma}
-	\label{lem:fixed_graph_testing}
-	Let $1 \leq \kappa \leq n$ be an integer.
-	\begin{enumerate}
-		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}$ is upper bounded
-		\begin{equation}
-		\label{eqn:graph_spectral_type_I_error}
-		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{1}{b^2}.
-		\end{equation}
-		\item \textbf{Type II error:} For any $b$ and $\beta$ such that
-		\begin{equation}
-		\label{eqn:fixed_graph_testing_critical_radius}
-		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{\beta^T L^s \beta}{n\lambda_{\kappa}^s}
-		\end{equation}
-		the Type II error of $\phi_{\spec}$ is upper bounded,
-		\begin{equation}
-		\label{eqn:graph_spectral_type_II_error}
-		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{3}{b}.
-		\end{equation}
-	\end{enumerate}
-\end{lemma}
+\subsection{Proof of Lemma~\ref{lem:fixed_graph_testing}}
+
+In this proof, we will drop all notational dependence on the graph $G$ for ease of reading.
 
@@
 
-\vspace{.2 in}
-
-\textit{Mean of $T_{\mathrm{spec}}$:} Using the notation $\dotp{v}{w} = \sum_{i = 1}^{n} v_iw_i$, we have
+\paragraph{Mean of $T_{\mathrm{spec}}$:} Using the notation $\dotp{v}{w} = \sum_{i = 1}^{n} v_iw_i$, we have
 \begin{align*}
@@
 
-\vspace{.2 in}
-
-\textit{Variance of $T_{\mathrm{spec}}$:}
+\paragraph{Variance of $T_{\mathrm{spec}}$:}
 We write $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvector $v_k$ as its $k$th column.
@@
 
-\vspace{.2 in}
-
-\textit{Proof of~\eqref{eqn:graph_spectral_type_I_error}:} By Chebyshev's inequality,
+\paragraph{Proof of~\eqref{eqn:graph_spectral_type_I_error}:} By Chebyshev's inequality,
 \begin{align*}
@@
 
-\vspace{.2 in}
-
-\textit{Proof of~\eqref{eqn:graph_spectral_type_II_error}:} For simplicity, we introduce the notation
+\paragraph{Proof of~\eqref{eqn:graph_spectral_type_II_error}:} For simplicity, we introduce the notation
 \begin{equation*}
@@
 
-\subsubsection{Step 2: Bounding neighborhood graph functionals}
-
-To make use of Lemma~\ref{lem:fixed_graph_testing} we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$.
+\subsection{Proof of Theorem~\ref{thm:sobolev_testing_rate_order1}}
+
+We want to apply Lemma~\ref{lem:fixed_graph_testing} to the case where $G = G_{n,r}$.
+In order to do this, we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$.
+In particular, we will show that the following statements each hold with probability at least $1 - c/b$ for sufficiently large $n$ (Here and in what follows, $c$ denotes a constant which is fixed in $n$ and $f$ and does not depend on $b$, but may depend on other fixed quantities such as $d$, $s$, etc.):
+\begin{enumerate}[label=(E\arabic*)]
+	\item 
+	\label{event:discrete_sobolev_norm_order1}
+	\textbf{Graph Sobolev norm:} For any $n^{-1/(2 + d)}\leq r \leq 1$,
+	\begin{equation}
+	\label{eqn:continuous_to_discrete_sobolev_norm_order1}
+	f^T L f \leq c \cdot b \cdot \norm{f}_{W^{1,2}(\Xset)}^2 \cdot n^{2} r^{d + 2} 
+	\end{equation}
+	\item 
+	\label{event:eigenvalue_tail_decay}
+	\textbf{Eigenvalue tail bound:} Let $t = n^{1/d}$, and let
+	\begin{equation*}
+	\overline{X} := \Bigl\{\frac{1}{t}(j_1,\ldots,j_d): k \in [t]^d\Bigr\}
+	\end{equation*} 
+	consist of $n$ evenly spaced grid points on $[0,1]^d$. The transportation distance between $X$ and $\ol{X}$ is
+	\begin{equation*}
+	d_{\infty}(X,\overline{X}) := \inf_{\pi} \sup\Bigl\{\norm{\pi(\ol{x}_i) - \ol{x}_i}: \ol{x}_i \in \ol{X}\Bigr\}
+	\end{equation*}
+	where the infimum is taken over all bijections $\pi: \ol{X} \to X$. 
+	There exists a constant $c$ such that for any $\kappa = 1,\ldots,n$ if $\max\{ 3 d_{\infty}(X,\ol{X}),n^{-1/d}\} \leq r \leq \kappa^{-2/(d(2 +d))}$, 
+	\begin{equation}
+	\label{eqn:eigenvalue_tail_bound}
+	\lambda_{\kappa} \geq c \cdot n r^{d + 2} \kappa^{2/d}
+	\end{equation}
+	Theorem~\ref{thm:slepcev_transport_distance} (Theorem 1.1 of \textcolor{red}{(Garcia Trillos and Slepcev)}) establishes that there exists a constant $c$ such that 
+	\begin{equation*}
+	d_{\infty}(X,\ol{X}) \leq c\biggl(\frac{\log n}{n}\biggr)^{1/d}
+	\end{equation*}
+	with probability $1  - O(1/n)$.
+Therefore in particular, the inequality~\eqref{eqn:eigenvalue_tail_bound} is satisfied for sufficiently large $n$ when $\kappa = n^{2d/(4 + d)}$ and $c(\log n/n)^{1/d} \leq r \leq n^{-4/((2+d)(4s + d))}$.
+	\item 
+	\label{event:l2_norm}
+	\textbf{Empirical norm of $f$:} There exists a constant $c_1$ such that if $\norm{f}_{\Leb^2(\Xset)} \geq c_1 \cdot b \cdot n^{-2/(4 + d)}$, and $d = 1,2$ or $3$,
+	\begin{equation}
+	\label{eqn:l2_to_empirical_norm}
+	\norm{f}_n^2 \geq \frac{1}{b} \cdot \norm{f}_{\Leb^2}^2
+	\end{equation}
+\end{enumerate}
+
+\subsubsection{Proof of \eqref{eqn:continuous_to_discrete_sobolev_norm}}
+
+The proof of~\eqref{eqn:continuous_to_discrete_sobolev_norm} follows from Lemma~\ref{lem:expected_first_order_seminorm}--which upper bounds the expected first order graph Sobolev seminorm-- and then an application of Markov's inequality.
+
+\subsubsection{Proof of \eqref{eqn:eigenvalue_tail_bound}}
+\label{subsubsec:eigenvalue_tail_bound_pf}
+
+We prove~\eqref{eqn:eigenvalue_tail_bound} by comparing $G_{n,r}$ to the tensor product of a $d$-dimensional lattice and a complete graph.
+The latter is a highly structured graph with known eigenvalues, which as we will see are sufficiently lower bounded for our purposes.
+
+Let $\wt{r} = r/(3(\sqrt{d} + 1)), M = (1/\wt{r})^d$ and $N = n\wt{r}^d$; assume without loss of generality that $M$ and $N$ are integers. Additionally for $m = M^{1/d}$ define
+\begin{equation*}
+\overline{Z} = \set{\frac{1}{m}(j_1,\ldots,j_d): j \in [m]^d}
+\end{equation*}
+to be the $M$ evenly spaced grid points over $[0,1]^d$.
+For a given $\overline{z}_j \in \overline{Z}$, we write $Q(z_j) = m^{-1}[j_1 - 1,j_1] \times \cdots \times m^{-1}[j_d - 1,j_d]$ for the cube of side length $1/m$ with $z_j$ at one corner.
+
+Consider the graph $H = (\overline{X}, E_H)$, where $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$ if
+\begin{equation*}
+\textrm{there exists}~\ol{z}_i, \ol{z}_j \in \ol{Z}~\textrm{such that}~\ol{x}_k \in Q(\ol{z}_i),~ \ol{x}_\ell \in Q(\ol{z}_j),~\textrm{and}~\norm{i - j}_1 \leq 1.
+\end{equation*}
+On the one hand $H \cong \ol{G}^M_d \otimes K_N$ where $\ol{G}^M_d$ is the $d$-dimensional lattice on $M$ nodes, and $K_N$ is the complete graph on $N$ nodes.
+On the other hand, we now show that when $\max\{3d_{\infty}(X,\ol{X}), n^{-1/d}\} \leq r$ then $G_{n,r} \succeq H$ as a result of the triangle inequality.
+If $(\ol{x}_k, \ol{x}_{\ell}) \in E(H)$, by definition there exist $\ol{z}_i, \ol{z}_j$ connected in $\ol{G}_d^M$ such that $\ol{x}_k \in Q(\ol{z}_i)$ and $\ol{x}_{\ell} \in Q(\ol{z}_j)$.
+This implies that $\ol{x}_k$ and $\ol{x}_{\ell}$ must themselves be close together, since
+\begin{equation*}
+\norm{\ol{x}_k - \ol{x}_{\ell}}_2 \leq \norm{\ol{x}_k - \ol{z}_{i}}_2 + \norm{\ol{z}_i - \ol{z}_j}_2 +  \norm{\ol{z}_{j} - \ol{x}_{\ell} }_2 \leq \wt{r}(1 + 2\sqrt{d}) = r/3.
+\end{equation*}
+Since we also assume $r/3 \geq d_{\infty}(X,\ol{X})$, another application of the triangle inequality gives
+\begin{equation*}
+\norm{\pi(\ol{x}_k) - \pi(\ol{x}_\ell)}_2 \leq \norm{\pi(\ol{x}_k) - \ol{x}_\ell}_2 + \norm{\ol{x}_k - \ol{x}_\ell}_2 \leq \norm{\ol{x}_{\ell} - \pi(\ol{x}_\ell)}_2 \leq r,
+\end{equation*}
+implying that $(\pi(\ol{x}_k), \pi(\ol{x}_{\ell})) \in E$ and consequently that $G_{n,r} \succeq \ol{G}^M_d \otimes K_N$.
+
+The eigenvalues of lattices and complete graphs are known to satisfy, respectively
+\begin{equation*}
+\lambda_k(\ol{G}^{M}_d) \geq \frac{k^{2/d}}{M^{2/d}}~\textrm{for $k = 0,\ldots,M - 1$},~~ \textrm{and}~\lambda_{j}(K_N) \geq N\1\{j > 0\}~\textrm{for $j = 0,\ldots,N-1$.}
+\end{equation*}
+and by standard facts regarding the eigenvalues of tensor product graphs, we have that the spectrum $\Lambda(H)$ satisfies
+\begin{equation*}
+\Lambda(H) = \set{N\lambda_k(\ol{G}^{M}_d) + M\lambda_j(K_N): \textrm{for $k = 0,\ldots,M - 1$ and $j = 0,\ldots,N-1$}}
+\end{equation*}
+For all $j = 1,\ldots,N-1$, we have that $M\lambda_j(K_N) = MN = n$. Therefore,
+\begin{align*}
+\lambda_{\kappa}(H) & \geq \{n \wedge N\lambda_{\kappa}(\ol{G}^{M}_d)\} \\
+& \geq \{n \wedge n\wt{r}^d\frac{\kappa^{2/d}}{M^{2/d}}\} \\
+& \geq \{n \wedge (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d}\} \\
+& \geq (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d},
+\end{align*}
+where the last inequality is satisfied since $r \leq \kappa^{-2/(d(d + 2))}$, completing the proof of~\eqref{eqn:eigenvalue_tail_bound}.
+
+\subsubsection{Proof of \eqref{eqn:l2_to_empirical_norm_order1}}
+The inequality~\eqref{eqn:l2_to_empirical_norm_order1} is implied by the more general Lemma~\ref{lem:empirical_norm_sobolev}, which we state in Section~\ref{subsec:sobolev_testing_rate_pf} and use to prove a similar inequality when $f \in W^{s,2}(\Xset)$ for general $s$.
+
+\subsubsection{Putting the pieces together}
+
+The final proof of Theorem~\ref{thm:sobolev_testing_rate_order1} proceeds by taking $s = 1$ and proceeding as in Section~\ref{subsubsec:sobolev_testing_rate_pf_conclusion}.
+
+\subsection{Proof of Theorem~\ref{thm:sobolev_testing_rate}}
+\label{subsec:sobolev_testing_rate_pf}
+
+We want to apply Lemma~\ref{lem:fixed_graph_testing} to the case where $G = G_{n,r}$.
+In order to do this, we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$.
 In particular, we will show that the following statements each hold with probability at least $1 - c/b$ for sufficiently large $n$ (Here and in what follows, $c$ denotes a constant which is fixed in $n$ and $f$ and does not depend on $b$, but may depend on other fixed quantities such as $d$, $s$, etc.):
@@
 	\label{event:discrete_sobolev_norm}
-	\textbf{Graph Sobolev norm:} For any $n^{-1/(2(s - 1) + d)}\leq r \leq 1$,
+	\textbf{Graph Sobolev norm:} For any $n^{-1/(2(s - 1) + d)}\leq r \leq \sigma/s$,
 	\begin{equation}
@@
 	\item 
-	\label{event:eigenvalue_tail_decay}
-	\textbf{Eigenvalue tail bound:} For any $\kappa = 1,\ldots,n$ and $r \leq \kappa^{-2/(d(2 +d))}$, 
-	\begin{equation}
-	\label{eqn:eigenvalue_tail_bound}
+	\label{event:eigenvalue_tail_decay_2}
+	\textbf{Eigenvalue tail bound:} There exists a constant $c$ such that for any $\kappa = 1,\ldots,n$, if $\max\{ 3 d_{\infty}(X,\ol{X}),n^{-1/d}\} \leq r \leq \kappa^{-2/(d(2 +d))}$ then,
+	\begin{equation}
+	\label{eqn:eigenvalue_tail_bound_2}
 	\lambda_{\kappa} \geq c \cdot n r^{d + 2} \kappa^{2/d}
 	\end{equation}
-	In particular, the inequality~\eqref{eqn:eigenvalue_tail_bound} is satisfied when $\kappa = n^{2d/(4s + d)}$ for any $r \leq n^{-4/((2+d)(4s + d))}$. 
+	In particular, the inequality~\eqref{eqn:eigenvalue_tail_bound_2} is satisfied for sufficiently large $n$ when $\kappa = n^{2d/(4s + d)}$ and $c(\log n/n)^{1/d} \leq r \leq n^{-4/((2+d)(4s + d))}$. 
 	\item 
-	\label{event:l2_norm}
-	\textbf{Empirical norm of $f$:} There exists a constant $c_1$ such that if $\norm{f}_{\Leb^2(\Xset)} \geq c_1 \cdot b \cdot n^{-ws/(4s + d)}$,
-	\begin{equation}
-	\label{eqn:l2_to_empirical_norm}
+	\label{event:l2_norm_order1}
+	\textbf{Empirical norm of $f$:} There exists a constant $c_1$ such that if $\norm{f}_{\Leb^2(\Xset)} \geq c_1 \cdot b \cdot n^{-2s/(4s + d)}$ and $4s > d$, then
+	\begin{equation}
+	\label{eqn:l2_to_empirical_norm_order1}
 	\norm{f}_n^2 \geq \frac{1}{b} \cdot \norm{f}_{\Leb^2}^2
@@
 
-\paragraph{Proof of~\eqref{eqn:continuous_to_discrete_sobolev_norm}:}
+\subsubsection{Proof of~\eqref{eqn:continuous_to_discrete_sobolev_norm}}
 
@@
 	\label{lem:roughness_functional_expectation_sobolev}
-	Let $\Xset$ be a Lipschitz domain, and let $s \geq 1$ be an integer.
-Suppose that $f \in W^{s,2}(\Xset)$, and further that $p \in C^{s-1}(\Xset;p_{\max})$ for some constant $p_{\max}$.
-Then for any $2$nd-order kernel $K$ and any $n^{-1/(2(s - 1) + d)} \leq r(n) \leq 1$, the expected graph Sobolev seminorm is upper bounded
+	Let $\Xset$ be a Lipschitz domain, let $s \geq 1$ be an integer, and let $0 < \sigma < 1$.
+Suppose that $f \in W_{\sigma}^{s,2}(\Xset)$, and further that $p \in C^{s-1}(\Xset;p_{\max})$ for some constant $p_{\max} < \infty$.
+Then for any $2$nd-order kernel $K$ and any $n^{-1/(2(s - 1) + d)} \leq r(n) < \sigma/s$, the expected graph Sobolev seminorm is upper bounded
 	\begin{equation}
 	\label{eqn:roughness_functional_expectation_sobolev}
-	\Ebb\bigl[f^T L^s f\bigr] \leq c \cdot \norm{f}_{W_d^{s,2}(\Xset)}^2 \cdot n^{s + 1}r^{s(d + 2)}
+	\Ebb\bigl[f^T L^s f\bigr] \leq c \cdot \norm{f}_{W^{s,2}(\Xset)}^2 \cdot n^{s + 1}r^{s(d + 2)}
 	\end{equation}
@@
 
+We note that the proof of Lemma~\ref{lem:roughness_functional_expectation_sobolev} is where we rely on the fact that $f$ and its derivatives are supported on $\Xset_{\sigma}$.
 The proof of Lemma~\ref{lem:roughness_functional_expectation_sobolev} is lengthy, and we defer it to Section~\ref{sec:technical_lemma_proofs}.
 
-\paragraph{Proof of~\eqref{eqn:eigenvalue_tail_bound}:}
-
-We prove~\eqref{eqn:eigenvalue_tail_bound} by comparing $G_{n,r}$ to the tensor product of a $d$-dimensional lattice and a complete graph.
-The latter is a highly structured graph with known eigenvalues, which as we will see are sufficiently lower bounded for our purposes.
-
-Let $\wt{r} = r/(3(\sqrt{d} + 1)), M = (1/\wt{r})^d, N = n\wt{r}^d$. Assume without loss of generality that $M$ and $N$ are integers. Additionally, for $t = n^{1/d}$ and $m = M^{1/d}$ let 
-\begin{equation*}
-\overline{X} = \set{\frac{1}{t}(k_1,\ldots,k_d): k \in [t]^d},~~ \overline{Z} = \set{\frac{1}{m}(j_1,\ldots,j_d): j \in [m]^d}.
-\end{equation*}
-For a given $\overline{z}_j \in \overline{Z}$, we write $Q(z_j) = m^{-1}[j_1 - 1,j_1] \times \cdots \times m^{-1}[j_d - 1,j_d]$ for the cube of side length $1/m$ with $z_j$ at one corner.
-
-Consider the graph $H = (\overline{X}, E_H)$, where $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$ if
-\begin{equation*}
-\textrm{there exists}~\ol{z}_i, \ol{z}_j \in \ol{Z}~\textrm{such that}~\ol{x}_k \in Q(\ol{z}_i),~ \ol{x}_\ell \in Q(\ol{z}_j),~\textrm{and}~\norm{i - j}_1 \leq 1.
-\end{equation*}
-On the one hand $H \cong \ol{G}^M_d \otimes K_N$ where $\ol{G}^M_d$ is the $d$-dimensional lattice on $M$ nodes, and $K_N$ is the complete graph on $N$ nodes.
-On the other hand, we now show that with high probability $G_{n,r} \succeq H$.
-If $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$, then there exist $\ol{z}_i, \ol{z}_j$ such that
-\begin{equation*}
-\norm{\ol{x}_k - \ol{x}_{\ell}}_2 \leq m^{-1} + \norm{\ol{x}_k - \ol{z}_{i}}_2 + \norm{\ol{x}_{\ell} - \ol{z}_{j}}_2 \leq \wt{r}(1 + \sqrt{d}) = r/3.
-\end{equation*}
-
-Assuming~\eqref{eqn:slepcev_transport_distance} holds, if $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$, then for sufficiently large $n$
-\begin{equation*}
-\norm{\pi(\ol{x}_k) - \pi(\ol{x}_\ell)}_2 \leq 2 c \left(\frac{\log n}{n}\right)^{1/d} + \frac{r}{3} \leq r,
-\end{equation*}
-implying that $(\pi(\ol{x}_k), \pi(\ol{x}_{\ell})) \in E$. Therefore, $G_{n,r} \succeq \ol{G}^M_d \otimes K_N$ whenever~\eqref{eqn:slepcev_transport_distance} holds.
-
-The eigenvalues of lattices and complete graphs are known to satisfy, respectively
-\begin{equation*}
-\lambda_k(\ol{G}^{M}_d) \geq \frac{k^{2/d}}{M^{2/d}}~\textrm{for $k = 0,\ldots,M - 1$},~~ \textrm{and}~\lambda_{j}(K_N) \geq N\1\{j > 0\}~\textrm{for $j = 0,\ldots,N-1$.}
-\end{equation*}
-and by standard facts regarding the eigenvalues of tensor product graphs, we have that the spectrum $\Lambda(H)$ satisfies
-\begin{equation*}
-\Lambda(H) = \set{N\lambda_k(\ol{G}^{M}_d) + M\lambda_j(K_N): \textrm{for $k = 0,\ldots,M - 1$ and $j = 0,\ldots,N-1$}}
-\end{equation*}
-For all $j = 1,\ldots,N-1$, we have that $M\lambda_j(K_N) = MN = n$. Therefore,
-\begin{align*}
-\lambda_{\kappa}(H) & \geq \{n \wedge N\lambda_{\kappa}(\ol{G}^{M}_d)\} \\
-& \geq \{n \wedge n\wt{r}^d\frac{\kappa^{2/d}}{M^{2/d}}\} \\
-& \geq \{n \wedge (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d}\} \\
-& \geq (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d},
-\end{align*}
-where the last inequality is satisfies since $r \leq \kappa^{-2/(d(d + 2))}$.
-
-Finally, Theorem~\ref{thm:slepcev_transport_distance} establishes that \eqref{eqn:slepcev_transport_distance} occurs with probability at least $1 - c/n$.
-Since $\lambda_{\kappa}(G_{n,r}) \geq \lambda_{\kappa}(H)$ when~\eqref{eqn:slepcev_transport_distance} is satisfied, this completes the proof of~\eqref{eqn:eigenvalue_tail_bound}.
-
-\paragraph{Proof of~\eqref{eqn:l2_to_empirical_norm}:}
+\subsubsection{Proof of~\eqref{eqn:eigenvalue_tail_bound}}
+See Section~\ref{subsubsec:eigenvalue_tail_bound_pf}.
+
+\subsubsection{Proof of~\eqref{eqn:l2_to_empirical_norm}}
 
@@
 The lower bound~\eqref{eqn:l2_to_empirical_norm} results from the more general Lemma~\ref{lem:empirical_norm_sobolev}, which can be verified by checking the various orderings of $2s/(4s + d)$, $s/d$ and $1/2$ whenever $4s < d$.
-\begin{proof}[Proof of Lemma~\ref{lem:empirical_norm_sobolev}]
+
+\paragraph{Proof of Lemma~\ref{lem:empirical_norm_sobolev}:}
+
 To prove~\eqref{eqn:l2_to_empirical_norm_sobolev} we will show
@@
 
-	\paragraph{Case 1: $2s > d$.}
+\textit{Case 1: $2s > d$.}
 When $2s > d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in C^{\gamma}(\overline{\Xset})$ for some $\gamma > 0$ which depends on $s$ and $d$, with the accompanying estimate
@@
 
-	\paragraph{Case 2: $2s < d$.}
+\textit{Case 2: $2s < d$.}
 When $2s < d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in \Leb^q(\Xset)$ for $q = 2d/(d - 2s)$, and moreover that
@@
 
-	\paragraph{Case 3: $2s = d$.}
+\textit{Case 3: $2s = d$.}
 Assume $f$ satisfies~\eqref{eqn:paley_zygmund_1} for a given $0 < a < 1$.
@@
 In particular the above holds for $q = 2/(1 - a)$ when $1/2 < a < 1$, and for any $q > 4$ when $0 < a < 1/2$. Using Lyapunov's inequality as in the previous case then implies the desired result.
-\end{proof}
-
-\subsubsection{Step 3: Putting the pieces together}
+
+\subsubsection{Putting the pieces together}
+\label{subsubsec:sobolev_testing_rate_pf_conclusion}
 
@@
 
-Now, we bound Type II error under the assumption $f \in W^{s,2}(\mathcal{X};R), p \in C^{s - 1}(\Xset;p_{\max})$ uniformly bounded away from $0$ and $\infty$ over $\Xset$, and 
+Now, we bound Type II error under the assumption $f \in W_{\sigma}^{s,2}(\mathcal{X};R), p \in C^{s - 1}(\Xset;p_{\max})$ uniformly bounded away from $0$ and $\infty$ over $\Xset$, and 
 \begin{equation}
 \label{eqn:critical_radius_1}
-\norm{f}_{\Leb^2}^2 \geq \epsilon^2 = c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4s + d)}\} \cdot n^{-4s/(4s + d)}.
+\norm{f}_{\Leb^2(\Xset)}^2 \geq \epsilon^2 = c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4s + d)}\} \cdot n^{-4s/(4s + d)}.
 \end{equation}
@@
 completing the proof of Theorem~\ref{thm:sobolev_testing_rate} upon proper choice of constants $c_1,c_2$ in that Theorem.
-
-\clearpage
 
@@
 \begin{cases*}
-O(r^{2s}) \cdot \norm{f}_{W_d^{s,2}(\Xset)}^2, & ~~\textrm{if $I = s + 1$} \\
-O(r^{2} \cdot r^{d(I - (s + 1))}) \cdot [f]_{W_d^{1,2}(\Xset)}^2, & ~~\textrm{if $I < s + 1$}~ 
+O(r^{2s}) \cdot \norm{f}_{W^{s,2}(\Xset)}^2, & ~~\textrm{if $I = s + 1$} \\
+O(r^{2} \cdot r^{d(I - (s + 1))}) \cdot [f]_{W^{1,2}(\Xset)}^2, & ~~\textrm{if $I < s + 1$}~ 
 \end{cases*}
@@
 
-Note that if $f$ is constant almost everywhere, the claim is immediate as $D_kf(x_i) = 0$ with probability one. Otherwise $[f]_{W^{1,2}(\Xset)} > 0$, which we shall assume in what follows.
+Note that if $f$ is constant almost everywhere in $\Xset$, the claim is immediate as $D_kf(x_i) = 0$ with probability one. Otherwise $[f]_{W^{1,2}(\Xset)} > 0$, which we shall assume in what follows.
 
 Let $\delta = \min\{r^{2s},1\}\cdot[f]_{W^{1,2}(\Xset)} > 0$.
-Our analysis will make heavy use of Taylor expansions, and we therefore would like to show that there exists some smooth $g \in C^{\infty}(\Rd)$ such that
-\begin{equation*}
-\Bigl|\Ebb[D_kf(x_i) D_{\ell} f(x_i)] - \Ebb[D_kg(x_i) D_\ell g(x_i)]\Bigr| <  \delta,
-\end{equation*}
-and additionally $[g]_{W^{\ell,2}(\Rd)} \leq c [f]_{W^{\ell,2}(\Rd)}$ for each $\ell = 0,\ldots,s$.
+Our analysis will make heavy use of Taylor expansions, and we therefore would like to show that there exists some $g \in C^{s}(\Xset) \cap W_{\sigma/2}^{s,2}(\Xset)$ such that
+\begin{equation*}
+\Bigl|\Ebb[D_kf(x_i) D_{\ell} f(x_i)] - \Ebb[D_kg(x_i) D_\ell g(x_i)]\Bigr| <  \delta, ~~\textrm{and}~~[g]_{W^{\ell,2}(\Rd)} \leq c [f]_{W^{\ell,2}(\Rd)}~\textrm{for each $\ell = 0,\ldots,s$.}
+\end{equation*}
 Then if \eqref{eqn:expected_difference_operators_1} and \eqref{eqn:expected_difference_operators_2} hold with respect to $g$, they hold (up to constants) with respect to $f$ as well. 
-
-To construct such a $g$, we first take an extension of $f$ be defined on all of $\Rd$, and then mollify.
-Since $\Xset$ is a Lipschitz domain, there exists an extension \textcolor{red}{citation} $\wt{f}$ of $f$ compactly supported on $\Rd$ such that $\wt{f} = f$ a.e.
-on $\Xset$, and $[\wt{f}]_{W^{\ell,2}(\Rd)} \leq c [f]_{W^{\ell,2}(\Xset)}$ for each $\ell = 0,\ldots,s$.
-Since $\wt{f} = f$ a.e.
-on $\Xset$, the expected difference operators satisfy $\Ebb[D_k\wt{f}(x_i) D_\ell \wt{f}(x_i)] = \Ebb[D_kf(x_i) D_\ell f(x_i)]$.
-
-Now, since $\wt{f} \in W^{s,2}(\Rd)$, there exists a sequence $(g_m) \subset C^{\infty}(\Rd)$ such that $\norm{g_m - \wt{f}}_{W^{s,2}(\Rd)} \to 0$. On the one hand, by the Cauchy-Schwarz inequality
-\begin{equation*}
-\abs{\Ebb\bigl[D_k\wt{f}(x_i) D_{\ell} \wt{f}(x_i)\bigr] - \Ebb\bigl[D_kg_m(x_i) D_\ell g_m(x_i)\bigr]} \leq \frac{c}{r^{sd}} \cdot \norm{f - g_m}_{\Leb^2(\Rd)}
+Note that since $f$ is supported on $\Xset_{\sigma}$, the extension
+\begin{equation*}
+\wt{f} := 
+\begin{cases*}
+f(x),~~ x \in \Xset_{\sigma} \\
+0,~~ x \in \Rd - \Xset_{\sigma}
+\end{cases*}
+\end{equation*}
+satisfies $\wt{f}(x) = f(x)$ for all $x \in \Xset$, and $\norm{\wt{f}}_{W^{s,2}(\Rd)} = \norm{f}_{W^{s,2}}(\Xset)$.
+
+To construct our function $g$, we consider as candidates the convolutions
+\begin{equation}
+g_m := \wt{f} \ast \eta_{1/m},~~\textrm{for each $m \in \mathbb{N}$.}
+\end{equation}
+where $\eta$ is the standard mollifier (for a definition see \textcolor{red}{Evans}.) We make the following observations:
+\begin{itemize}
+	\item Since $\eta$ is a smooth function, $g_m \in C^{\infty}(\Rd)$ and therefore $g_m \in C^s(\Xset)$. 
+	\item For all sufficiently large $m$ and all multiindices $\alpha$,
+	\begin{equation*}
+	g_m^{(\alpha)}(x) = 0~~\textrm{if $x \in \Reals^d - X_{\sigma/2}$}.
+	\end{equation*}
+	and along with the previous observation this implies $g_m \in W_{\sigma/2}^{s,2}(\Xset)$. 
+	\item By Theorem~1 of \textcolor{red}{Evans 5.3.1}, $\norm{g_m - \wt{f}}_{W_{\mathrm{loc}}^{s,2}(\Rd)} \to 0$ as $m \to \infty$, which implies  $\norm{g_m - f}_{W^{s,2}(\Xset)} \to 0$.
+\end{itemize}
+By the Cauchy-Schwarz inequality
+\begin{equation*}
+\abs{\Ebb\bigl[D_kf(x_i) D_{\ell} f(x_i)\bigr] - \Ebb\bigl[D_kg_m(x_i) D_\ell g_m(x_i)\bigr]} \leq \frac{c}{r^{sd}} \cdot \norm{f - g_m}_{\Leb^2(\Xset)}
 \end{equation*}
 and taking $m$ to be sufficiently large, we can make the right hand side less than $\delta$.
-On the other hand, since $\norm{g_m - \wt{f}}_{W^{s,2}(\Rd)} \to 0$, there exists $m$ sufficiently large such that $[g_m]_{W^{s,2}(\Rd)}$ is at most say $2 [\wt{f}]_{W^{s,2}(\Rd)}$.
+On the other hand, since $\norm{g_m - f}_{W^{s,2}(\Xset)} \to 0$, there exists $m$ sufficiently large such that $[g_m]_{W^{\ell,2}(\Xset)}$ is at most say $2 [f]_{W^{\ell,2}(\Xset)}$ for each $\ell = 0,\ldots,s$.
 We therefore take $g = g_m$ for $m$ large enough to satisfy both conditions.
@@
 & = O(r^{(\abs{k \cup \ell \cup i} - 3)d}) \cdot \Ebb\left[\bigl(d_{i}g(x_j)\bigr)^2K_r(x_i,x_j)\right] \\
-& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot [g]_{W^{1,2}(\Rd)}^2
+& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot [g]_{W^{1,2}(\Xset)}^2
 \end{align*}
@@
 
-By Lemma~\ref{lem:leading_term_sobolev} there exists some $f_s \in \Leb^2(\Rd)$ which satisfies $\norm{f_s}_{\Leb^2(\Rd)} \leq c \norm{g}_{W^{s,2}(\Rd)}$ such that
-\begin{equation*}
-\Ebb\Bigl[\bigl(D_kg(x)\bigr)^2\Bigr] = r^{2s} \cdot \norm{f_s}_{\Leb^2(\Rd)}.
-\end{equation*} Therefore, by the law of iterated expectation along with this Lemma, 
-\begin{align*}
-\Ebb\bigl[D_kg(x_i)D_kg(x_j)\bigr] = \Ebb\Bigl[\bigl(\Ebb[D_kg(x_i)|x_i]\bigr)^2\Bigr] = r^{2s} \cdot \Ebb\Bigl[\bigl(f_s(x_i)\bigr)^2\Bigr] = r^{2s} \cdot \norm{g}_{W^{s,2}(\Rd)},
+By Lemma~\ref{lem:leading_term_sobolev} there exists some $f_s \in \Leb^2(\Xset)$ which satisfies $\norm{f_s}_{\Leb^2(\Xset)} \leq c \norm{g}_{W^{s,2}(\Xset)}$ such that
+\begin{equation*}
+\Ebb\Bigl[\bigl(D_kg(x)\bigr)^2\Bigr] = r^{2s} \cdot \norm{f_s}_{\Leb^2(\Xset)}.
+\end{equation*} Therefore, by the law of iterated expectation 
+\begin{align*}
+\Ebb\bigl[D_kg(x_i)D_kg(x_j)\bigr] = \Ebb\Bigl[\bigl(\Ebb[D_kg(x_i)|x_i]\bigr)^2\Bigr] = r^{2s} \cdot \Ebb\Bigl[\bigl(f_s(x_i)\bigr)^2\Bigr] \leq r^{2s} \cdot c \norm{g}_{W^{s,2}(\Xset)},
 \end{align*}
@@
 By the law of iterated expectation, we have
-\begin{align*}
-\Ebb[d_iD_kg_m(x_j)d_iD_{\ell}g_m(x_j)K_r(x_i,x_j)] & = \Ebb\biggl[\Bigl(d_i\bigl(\Ebb(D_kf)\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \\
-& = \Ebb\biggl[\Bigl(d_i\bigl(r^{s - 1}\cdot f_{s - 1} + r^sf_s(x_j)\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr].
-\end{align*}
-where the latter equality follows from Lemma~\ref{lem:leading_term_sobolev}. Here $f_{s -1}$ and $f_s$ satisfy the conclusions of that Lemma, namely that $f_{s - 1}$ and $f_s \in C^{\infty}(\Rd)$, and
-\begin{equation*}
-\norm{f_{s - 1}}_{W^{1,2}(\Rd)}, \norm{f_s}_{\Leb^2(\Rd)} \leq c \norm{g}_{W^{s,2}(\Rd)}.
-\end{equation*}
-By the linearity of the difference operator $d_i$, we have
+\begin{align}
+\Ebb\biggl[\Bigl(d_iD_kg_m(x_j)\Bigr) \Bigl(d_iD_{\ell}g_m(x_j)\Bigr) K_r(x_i,x_j)\biggr] & = \Ebb\biggl[\Bigl(d_i\bigl(\Ebb[D_kf|x_i,x_j]\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \nonumber \\
+& = \Ebb\biggl[\Bigl(d_i\bigl(r^{s - 1}\cdot f_{s - 1} + r^sf_s(x_j)\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr]. \label{eqn:expected_difference_operators_sobolev_pf1}
+\end{align}
+where the latter equality follows from Lemma~\ref{lem:leading_term_sobolev} and $f_{s -1}$ and $f_s$ satisfy the conclusions of that Lemma, namely that $f_{s - 1} \in C^1(\Xset) \cap W^{1,2}(\Xset)$ and $f_s \in C^{0}(\Xset) \cap \Leb^2(\Xset)$, and
+\begin{equation*}
+\norm{f_{s - 1}}_{W^{1,2}(\Xset)}, \norm{f_s}_{\Leb^2(\Xset)} \leq c \norm{g}_{W^{s,2}(\Xset)}.
+\end{equation*}
+Applying these estimates inside \eqref{eqn:expected_difference_operators_sobolev_pf1}, we obtain
 \begin{align*}
@@
 \end{align*}
-As a result of~\eqref{eqn:transport_distance} we obtain the following bounds on $M(\gamma)$ and $b(\gamma)$:
+As a result of~\eqref{eqn:slepcev_transport_distance} we obtain the following bounds on $M(\gamma)$ and $b(\gamma)$:
 \begin{align}
@@
 
+For Lemmas~\ref{lem:expected_first_order_seminorm} - \ref{lem:expected_first_order_seminorm_3}, we will assume that $K$ is a kernel function compactly supported on $B(0,1)$ and upper bound $K(x) \leq K_{\max}$.
+
 \begin{lemma}
 	\label{lem:expected_first_order_seminorm}
-	Suppose $g \in C^{\infty}(\Rd)$, that $\abs{p(x)} \leq p_{\max}$ for all $x \in \Rd$, and that $K$ is a $2$nd order kernel. Then
-	\begin{equation*}
-	\Ebb[(g(x_j) - g(x_i))^2K_r(x_i,x_j)] \leq c K_{\max} p_{\max}^2 r^2 [g]_{W^{1,2}(\Rd)}^2
+	Suppose $g \in C^{1}(\Xset)$ for $\Xset$ a Lipschitz domain and that $\abs{p(x)} \leq p_{\max}$ for all $x \in \Xset$. Then
+	\begin{equation*}
+	\Ebb\Bigl[(g(x_j) - g(x_i))^2K_r(x_i,x_j)\Bigr] \leq c K_{\max} p_{\max}^2 r^2 [g]_{W^{1,2}(\Xset)}^2
 	\end{equation*}
@@
 \begin{proof}
+	Since $\Xset$ is a Lipschitz domain, we may take $h \in C^1(\Rd)$ to be an extension of $g$ such that $h = g$ a.e. on $\Xset$, and additionally
+	\begin{equation*}
+	[h]_{W^{1,2}(\Rd)} \leq c[g]_{W^{1,2}(\Xset)}
+	\end{equation*}
+	Since $h = g$ a.e on $\Xset$, the expectations satisfy
+	\begin{equation*}
+	\Ebb\Bigl[(g(x_j) - g(x_i))^2K_r(x_i,x_j)\Bigr] = \Ebb\Bigl[(h(x_j) - h(x_i))^2K_r(x_i,x_j)\Bigr].
+	\end{equation*}
 	By the fundamental theorem of calculus we have for any $y,x \in \Rd$,
-	\begin{equation*}
-	g(y) - g(x) = \int_{0}^{1} \frac{d}{dt}\bigl[g(x + t(y - x))\bigr] \,dt = \int_{0}^{1} \dotp{\nabla(g(x + t(y - x)))}{y - x} \,dt
-	\end{equation*}
-	By the upper bound on $p$, we obtain
-	\begin{align*}
-	\Ebb[(g(x_j) - g(x_i))^2K_r(x_i,x_j)] & \leq p_{\max}^2 \int_{\Rd} \int_{\Rd} (g(y) - g(x))^2 K_r(y,x) \,dy \,dx\\
-	& = p_{\max}^2 \int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \dotp{\nabla(g(x + t(y - x)))}{y - x} \,dt\right)^2 K_r(y,x) \,dy \,dx \\
-	& \overset{(i)}{\leq} p_{\max}^2 \int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \norm{\nabla(g(x + t(y - x)))}\norm{y - x} \,dt\right)^2 K_r(y,x) \,dy \,dx \\
-	& \overset{(ii)}{\leq} p_{\max}^2 r^2\int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \norm{\nabla(g(x + t(y - x)))}\,dt\right)^2 K_r(y,x) \,dy \,dx \\
-	& \overset{(iii)}{\leq} p_{\max}^2 r^2\int_{\Rd} \int_{\Rd} \int_{0}^{1} \norm{\nabla(g(x + t(y - x)))}^2 \,dt K_r(y,x) \,dy \,dx \\
-	& \overset{(iv)}{\leq} p_{\max}^2 K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(x,r)} \norm{\nabla(g(x + t(y - x)))}^2 \,dy \,dt \,dx \\
-	& \overset{(v)}{\leq}  p_{\max}^2 K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(0,r)} \norm{\nabla(g(x + z))}^2  \,dz \,dt \,dx
+	\begin{equation}
+	\label{eqn:expected_first_order_seminorm_pf1}
+	h(y) - h(x) = \int_{0}^{1} \frac{d}{dt}\bigl[h(x + t(y - x))\bigr] \,dt = \int_{0}^{1} \dotp{\nabla(h(x + t(y - x)))}{y - x} \,dt
+	\end{equation}
+	where the integral is well-defined as $\nabla h(z)$ exists almost everywhere since $h \in C^1(\Rd)$. We now perform some standard calculus:
+	\begin{align*}
+	\Ebb[(h(x_j) - h(x_i))^2K_r(x_i,x_j)] & \leq p_{\max}^2 \int_{\Rd} \int_{\Rd} (h(y) - h(x))^2 K_r(y,x) \,dy \,dx\\
+	& = p_{\max}^2 \int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \dotp{\nabla(h(x + t(y - x)))}{y - x} \,dt\right)^2 K_r(y,x) \,dy \,dx \\
+	& \overset{(i)}{\leq} p_{\max}^2 \int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \norm{\nabla(h(x + t(y - x)))}\norm{y - x} \,dt\right)^2 K_r(y,x) \,dy \,dx \\
+	& \overset{(ii)}{\leq} p_{\max}^2 r^2\int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \norm{\nabla(h(x + t(y - x)))}\,dt\right)^2 K_r(y,x) \,dy \,dx \\
+	& \overset{(iii)}{\leq} p_{\max}^2 r^2\int_{\Rd} \int_{\Rd} \int_{0}^{1} \norm{\nabla(h(x + t(y - x)))}^2 \,dt K_r(y,x) \,dy \,dx \\
+	& \overset{(iv)}{\leq} p_{\max}^2 K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(x,r)} \norm{\nabla(h(x + t(y - x)))}^2 \,dy \,dt \,dx \\
+	& \overset{(v)}{\leq}  p_{\max}^2 K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(0,r)} \norm{\nabla(h(x + z))}^2  \,dz \,dt \,dx
 	\end{align*}
@@
 	\begin{align*}
-	K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(0,r)} \norm{\nabla(g(x + z))}^2  \,dz \,dt \,dx & = r^{2 - d}\int_{B(0,r)} \int_{0}^{1} \int_{\Rd} \norm{\nabla(g(x + z))}^2  \,dz \,dt \,dx \\
-	& = K_{\max} r^2 [g]_{W_d^{1,2}(\Rd)}.
+	K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(0,r)} \norm{\nabla(h(x + z))}^2  \,dz \,dt \,dx & = r^{2 - d}\int_{B(0,r)} \int_{0}^{1} \int_{\Rd} \norm{\nabla(h(x + z))}^2  \,dz \,dt \,dx \\
+	& = K_{\max} r^2 [h]_{W_d^{1,2}(\Rd)}.
 	\end{align*}
@@
 	\label{lem:expected_first_order_seminorm_2}
-	Suppose $g \in C^{\infty}(\Rd)$, that $\abs{p(x)} \leq p_{\max}$ for all $x \in \Rd$, and that $K$ is a $2$nd order kernel. Then
-	\begin{equation*}
-	\Ebb\Bigl[\abs{D_ig(x_h)}\cdot\abs{D_ig(x_j)} \Bigr] \leq c K_{\max} p_{\max}^2 r^2 [g]_{W^{1,2}(\Rd)}^2
+	Suppose $f \in C^{1}(\Xset)$ for $\Xset$ a Lipschitz domain and that $\abs{p(x)} \leq p_{\max}$ for all $x \in \Xset$. Then
+	\begin{equation*}
+	\Ebb\Bigl[\abs{D_if(x_h)}\cdot\abs{D_if(x_j)} \Bigr] \leq c K_{\max}^3 p_{\max}^3 r^2 [f]_{W^{1,2}(\Xset)}^2
 	\end{equation*}
@@
 \begin{proof}
+	Since $\Xset$ is a Lipschitz domain, we may take $g \in C^1(\Rd)$ to be an extension of $f$ such that $g = f$ a.e. on $\Xset$, and additionally
+	\begin{equation*}
+	[g]_{W^{1,2}(\Rd)} \leq c[f]_{W^{1,2}(\Xset)}
+	\end{equation*}
+	Since $g = f$ a.e on $\Xset$, the expectations satisfy
+	\begin{equation*}
+	\Ebb\Bigl[\abs{D_if(x_h)}\cdot\abs{D_if(x_j)} \Bigr] =\Ebb\Bigl[\abs{D_ig(x_h)}\cdot\abs{D_ig(x_j)} \Bigr]
+	\end{equation*}
+
 	We rewrite $\Ebb\bigl[\abs{D_{i}g(x_j)} \cdot \abs{D_ig(x_h)} \bigr]$ as follows,
@@
 	\end{align*}
-	Then we obtain
+	Applying~\eqref{eqn:expected_first_order_seminorm_pf1} inside the integral gives
 	\begin{align*}
@@
 	\begin{equation*}
-	p_{\max}^3 \int_{\Xset} \left[\int_{\Xset} \abs{g(z) - g(x)} K_r(z,x) \,dx\right]^2 \,dz \leq c\cdot p_{\max}^3 r^2 K_{\max}^3 [f]_{W_d^{1,2}(\Rd)}^2 = O(r^2)\cdot [f]_{W_d^{1,2}(\Xset)}^2.
+	p_{\max}^3 \int_{\Xset} \left[\int_{\Xset} \abs{g(z) - g(x)} K_r(z,x) \,dx\right]^2 \,dz \leq c\cdot p_{\max}^3 r^2 K_{\max}^3 [f]_{W_d^{1,2}(\Rd)}^2.
 	\end{equation*}
@@
 	\label{lem:expected_first_order_seminorm_3}
-	Suppose $g \in C^{\infty}(\Rd)$, that $\abs{p(x)} \leq p_{\max}$ for all $x \in \Rd$, and that $K$ is a $2$nd order kernel. Then
-	\begin{equation*}
-	\Ebb\Bigl[\abs{D_{k_q}g(x_i)}\cdot{\abs{D_{\ell_q}g(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr] \leq c K_{\max} p_{\max}^2 r^{2 + d} [g]_{W^{1,2}(\Rd)}^2
+	Suppose $f \in C^{1}(\Xset)$ and that $\abs{p(x)} \leq p_{\max}$ for all $x \in \Xset$. Then for any distinct $i,j,k,\ell$ each in $[n]$,
+	\begin{equation*}
+	\Ebb\Bigl[\abs{D_{k}f(x_i)}\cdot{\abs{D_{\ell}f(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr] \leq c K_{\max} p_{\max}^2 r^{2 + d} [f]_{W^{1,2}(\Rd)}^2
 	\end{equation*}
@@
 \begin{proof}
+	Since $\Xset$ is a Lipschitz domain, we may take $g \in C^1(\Rd)$ to be an extension of $f$ such that $g = f$ a.e. on $\Xset$, and additionally
+	\begin{equation*}
+	[g]_{W^{1,2}(\Rd)} \leq c[f]_{W^{1,2}(\Xset)}
+	\end{equation*}
+	Since $g = f$ a.e on $\Xset$, the expectations satisfy
+	\begin{equation*}
+	\Ebb\Bigl[\abs{D_{k}f(x_i)}\cdot{\abs{D_{\ell}f(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr] = 	\Ebb\Bigl[\abs{D_{k}g(x_i)}\cdot{\abs{D_{\ell}g(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr].
+	\end{equation*}
 	We rewrite the expectation as an integral,
 	\begin{align*}
-	\Ebb\Bigl[& \abs{D_{k_q}g(x_i)}\cdot{\abs{D_{\ell_q}g(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr] \\
+	\Ebb\Bigl[& \abs{D_{k}g(x_i)}\cdot{\abs{D_{\ell}g(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr] \\
 	& \leq p_{\max}^4 \int_{\Xset^4} \abs{g(x) - g(y)} \cdot \abs{g(u) - g(v)} \cdot  K_r(x,y) K_r(u,v) \1\{\norm{y - v} \leq (2q + 1)r\} \,dy \,dx \,du \,dv
@@
 	\label{lem:remainder_term}
-	Suppose $f \in \Leb^2(\Rd)$, and $h:\Rd \times \Rd \times [0,1] \to \Reals$ is uniformly bounded.
-Then, the function $g(x) = \int_{0}^{1} \int_{B(0,1)} f(x + aty) h(y,x, t) \,dy \,dt$ also belongs to $\Leb^2(\Rd)$, with norm
-	\begin{equation*}
-	\norm{g}_{\Leb^2(\Rd)} \leq \nu_d \cdot \norm{f}_{\Leb^2(\Rd)} \cdot \norm{h}_{\infty}
+	Suppose that $f \in \Leb^2(U)$ for some open domain $U$, and that $h:U \times U \times [0,1] \to \Reals$ is uniformly bounded.
+Then, the function $g(x) = \int_{0}^{1} \int_{B(0,1)} f(x + aty) h(y,x, t) \,dy \,dt$ also belongs to $\Leb^2(U)$, with norm
+	\begin{equation*}
+	\norm{g}_{\Leb^2(U)} \leq \nu_d \cdot \norm{f}_{\Leb^2(U)} \cdot \norm{h}_{\infty}
 	\end{equation*}
@@
 	\begin{align*}
-	\norm{g}_{\Leb^2(\Rd)}^2 & = \int_{\Rd} \left(\int_{0}^{1} \int_{B(0,1)} f(x + aty) h(y,x,t) \,dt \,dy \right)^2 \,dx \\
-	& \leq \norm{h}_{\infty}^2 \int_{\Rd} \left(\int_{0}^{1} \int_{B(0,1)} f(x + aty) \,dt \,dy \right)^2 \,dx \\
-	& \leq \nu_d^2 \norm{h}_{\infty}^2 \int_{\Rd} \int_{0}^{1} \frac{1}{\nu_d}\int_{B(0,1)} f^2(x + aty) \,dt \,dy \,dx \tag{Jensen's inequality} \\
-	& = \nu_d^2 \norm{h}_{\infty}^2 \int_{0}^{1} \int_{B(0,1)} \frac{1}{\nu_d}\int_{\Rd}f^2(x + aty) \,dt \,dy \,dx \tag{Fubini's theorem} \\
-	& = \nu_d^2 \norm{h}_{\infty}^2 \norm{f}_{\Leb^2(\Rd)}^2.
+	\norm{g}_{\Leb^2(\Rd)}^2 & = \int_{U} \left(\int_{0}^{1} \int_{B(0,1)} f(x + aty) h(y,x,t) \,dt \,dy \right)^2 \,dx \\
+	& \leq \norm{h}_{\infty}^2 \int_{U} \left(\int_{0}^{1} \int_{B(0,1)} f(x + aty) \,dt \,dy \right)^2 \,dx \\
+	& \leq \nu_d^2 \norm{h}_{\infty}^2 \int_{U} \int_{0}^{1} \frac{1}{\nu_d}\int_{B(0,1)} f^2(x + aty) \,dt \,dy \,dx \tag{Jensen's inequality} \\
+	& = \nu_d^2 \norm{h}_{\infty}^2 \int_{0}^{1} \int_{B(0,1)} \frac{1}{\nu_d}\int_{U}f^2(x + aty) \,dt \,dy \,dx \tag{Fubini's theorem} \\
+	& = \nu_d^2 \norm{h}_{\infty}^2 \norm{f}_{\Leb^2(U)}^2.
 	\end{align*}
@@
 	\label{lem:leading_term_sobolev}
-	Let $k \in [n]^q$ for some $q \geq 1$, and let $K_r$ be a second order kernel.
-For $s \geq 0$ an integer, let $p$ be a density satisfying $p \in C^{s - 1}(\Rd;p_{\max})$ for some $p_{\max} > 0$ if $s \geq 1$, and otherwise $p \in C^{0}(\Rd;p_{\max})$.
-Then for any $f \in C^{\infty}(\Rd)$, there exist functions $f_{\ell}, \ell = 2q,\ldots,s$ satisfying $f_{\ell} \in C^{\infty}(\Rd)$ and
-	\begin{equation*}
-	\norm{f_{\ell}}_{W^{s - \ell,2}(\Rd)} \leq c \norm{f}_{W^{s,2}(\Rd)}
-	\end{equation*}
-	for some constant $c$ which depends only on $d$, $\Xset$ and $p_{\max}$ such that
+	Let $k \in [n]^q$ for some $q \geq 1$, let $K_r$ be a second order kernel, and suppose $f \in C^{s}(\Xset) \cap W_{\sigma}^{s,2}(\Xset)$ for some $s \in \mathbb{N}$ and $\sigma > 0$.
+Let $p$ be a density satisfying $p \in C^{0}(\Xset;p_{\max})$ if $s = 0$, and otherwise $p \in C^{s - 1}(\Xset;p_{\max})$ for some $p_{\max} > 0$.
+For any $qr < \sigma$, there exist functions $f_{\ell}$ for $\ell = 2q,\ldots,s - 1$ and $f_s$ satisfying $f_{\ell} \in C^{s - \ell}(\Xset) \cap W_{\sigma - qr}^{s - \ell,2}(\Xset)$ and
+	\begin{equation*}
+	\norm{f_{\ell}}_{W^{s - \ell,2}(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}
+	\end{equation*}
+	for some constant $c$ which depends only on $\sigma$, $d$, $\Xset$ and $p_{\max}$ such that
 	\begin{equation}
@@
 	\end{equation}
+	for any $x \in \Xset$.
 \end{lemma}
@@
 	We proceed by induction on $q$.
-Note that without loss of generality, we will assume $p \in C^{\infty}(\Rd)$, i.e.
-$p$ is smooth.
-Otherwise we can simply mollify $p$, with neglible change to the Holder norm $\norm{p}_{C^{s-1}(\Xset)}$ or the expectation $\Ebb[D_kf(x)]$.
 
 	\paragraph{Base case.}
-	We begin with the base case of $q = 1$. When $s = 0$, we need to show that $\Ebb[D_kf(x)] = f_s(x)$ for some $\norm{f_s}_{\Leb^2(\Rd)} \leq c\norm{f}_{W^{s,2}(\Rd)}$. But
-	\begin{align*}
-	\Ebb\Bigl[D_kf(x)\Bigr] & = \int f(z) K_r(x,z) p(z) \,dz  - f(x) \\
-	& = \int f(yr + x) K(\norm{y}) p(yr + z) \,dy - f(x).
-	\end{align*}
-	and since
-	\begin{equation*}
-	\biggl\|\int f(yr + x) K(\norm{y}) p(yr + x) \,dy\biggr\|_{\Leb^2(\Rd)} \leq p_{\max} K_{\max} \norm{f}_{\Leb^2(\Rd)}
+	We begin with the base case of $q = 1$.
+When $s = 0$, we need to show that $\Ebb[D_kf(x)] = f_s(x)$ for some $f_s \in C^{0}(\Xset)$ satisfying $\mathrm{supp}(f_s) \subset X_{\sigma - r}$ and $\norm{f_s}_{\Leb^2(\Xset)} \leq c\norm{f}_{W^{s,2}(\Xset)}$.
+But
+	\begin{align*}
+	\Ebb\Bigl[D_kf(x)\Bigr] & = \int f(z) K_r(x,z) p(z) \,dz  - f(x)\Ebb\bigl[K_r(x_k,x)\bigr] \\
+	& = \int f(yr + x) K\bigl(\norm{y}\bigr) p(yr + z) \,dy - f(x)\Ebb\bigl[K_r(x_k,x)\bigr],
+	\end{align*}
+	whence the boundedness $f_s \in C^{0}(\Xset)$ follows from the boundedness of $f, K$ and $p$.
+We now analyze each term in the above difference.
+First, we have that $f(x)\Ebb[K_r(x_k,x)] = 0$ unless $x \in \Xset_{\sigma}$, and
+	\begin{equation*}
+	\norm{f\Ebb[K_r(x_k,\cdot)]}_{\Leb^2(\Xset)} \leq p_{\max} \norm{f}_{\Leb^2(\Xset)}.
+	\end{equation*}
+	Moreover, since $K$ is compactly supported on $B(0,1)$, the integral $\int f(yr + x) K(\norm{y}) p(yr + x) \,dy = 0$ unless $x \in \Xset_{\sigma - r}$. Since
+	\begin{equation*}
+	\biggl\|\int f(yr + x) K(\norm{y}) p(yr + x) \,dy\biggr\|_{\Leb^2(\Xset)} \leq p_{\max} K_{\max} \norm{f}_{\Leb^2(\Xset)},
 	\end{equation*}
@@
 
-	Now, when $s \geq 1$ since $f$ is smooth it admits a Taylor expansion of the following form for all $x,z \in \Rd$:
+	Now, when $s \geq 1$ since $f \in C^{s}(\Xset)$ it admits a Taylor expansion of the following form for all $x,z \in \Xset$:
 	\begin{equation*}
@@
 	\end{equation*}
-	where $f^{(\alpha)} \in C^{\infty}(\Rd)$ additionally satisfies
-	\begin{equation*}
-	\norm{f^{(\alpha)}}_{W^{s - \abs{\alpha},2}(\Rd)} \leq \norm{f}_{W^{s,2}(\Rd)}
+	$f^{(\alpha)} \in C^{s - \abs{\alpha}}(\Xset)$ additionally satisfies
+	\begin{equation*}
+	\norm{f^{(\alpha)}}_{W^{s - \abs{\alpha},2}(\Xset)} \leq \norm{f}_{W^{s,2}(\Xset)} ~~\textrm{and}~~ \mathrm{supp}\bigl(f^{(\alpha)}\bigr) \subset {\Xset}_{\sigma}.
 	\end{equation*}
@@
 	\label{eqn:leading_term_sobolev_pf1}
-	\Ebb(D_kf(x)) = \sum_{1 < \abs{\alpha} < s} \frac{f^{(\alpha)}(x)}{\alpha!} E_{\alpha,x} + \frac{1}{(s - 1)!}\sum_{\abs{\alpha} = s} \int_{0}^{1}(1 - t)^{s - 1} E_{\alpha,x,t}(f)  \,dt 
+	\Ebb(D_kf(x)) = \sum_{1 \leq \abs{\alpha} < s} \frac{f^{(\alpha)}(x)}{\alpha!} E_{\alpha,x} + \frac{1}{(s - 1)!}\sum_{\abs{\alpha} = s} \int_{0}^{1}(1 - t)^{s - 1} E_{\alpha,x,t}(f)  \,dt 
 	\end{equation}
@@
 	\begin{align*}
-	E_{\alpha,x,t}(f) & = \frac{1}{r^d}\int_{\Rd} f^{(\alpha)}\bigl(x + t(z - x)\bigr) (z - x)^{\alpha}  K\Bigl(\frac{\norm{z - x}}{r}\Bigr) p(z) \,dz \\
-	& = r^s \int_{\Rd} y^{\alpha} f^{(\alpha)}\bigl(x + tyr) K(\norm{y}) p(yr + x) \,dy,
+	E_{\alpha,x,t}(f) & = \frac{1}{r^d}\int_{\Rd} f^{(\alpha)}\bigl(x + t(z - x)\bigr) (z - x)^{\alpha}  K\biggl(\frac{\norm{z - x}}{r}\biggr) p(z) \,dz \\
+	& = r^s \int_{\Rd} y^{\alpha} f^{(\alpha)}\bigl(x + tyr) K\bigl(\norm{y}\bigr) p(yr + x) \,dy
 	\end{align*}
@@
 	\begin{align*}
-	\frac{1}{(s - 1)!}\sum_{\abs{\alpha} = s} \int_{0}^{1}(1 - t)^{s - 1} E_{\alpha,x,t}(f)  \,dt & = \frac{r^s}{(s - 1)!}\sum_{\abs{\alpha} = s} \int_{0}^{1} \int_{\Rd} (1 - t)^{s - 1}  y^{\alpha} f^{(\alpha)}\bigl(x + tyr) K(\norm{y}) p(yr + x) \,dy \,dt \\
-	& =: r^s g_s(x) ~~\textrm{for $g_s \in C^{\infty}(\Rd)$} 
-	\end{align*}
-	where $\norm{g_s}_{\Leb^2(\Rd)} \leq c \norm{f}_{W^{s,2}(\Rd)}$ follows from Lemma~\ref{lem:remainder_term}.
-When $s = 1$, there are no $1 < \abs{\alpha} < s$, and so only the remainder term in \eqref{eqn:leading_term_sobolev_pf1} is non-zero; thus we have shown~\eqref{eqn:leading_term_sobolev} when $q = 1$ and $s = 1$.
-
-	When $s \geq 2$, we can analyze $E_{\alpha,x}$ using a Taylor expansion of $p$. For any $x,z \in \Rd$, we have
+	\frac{1}{(s - 1)!}\sum_{\abs{\alpha} = s} \int_{0}^{1}(1 - t)^{s - 1} E_{\alpha,x,t}(f)  \,dt & = \frac{r^s}{(s - 1)!}\sum_{\abs{\alpha} = s} \int_{0}^{1} \int_{\Rd} (1 - t)^{s - 1}  y^{\alpha} f^{(\alpha)}\bigl(x + tyr) K\bigl(\norm{y}\bigr) p(yr + x) \,dy \,dt \\
+	& =: r^s g_s(x) ~~\textrm{for $g_s \in C^{0}(\Xset) \cap \Leb_{\sigma - r}^2(\Xset)$},
+	\end{align*}
+	where additionally $\norm{g_s}_{\Leb^2(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}$ by Lemma~\ref{lem:remainder_term}.
+When $s = 1$, there are no multiindices $1 < \abs{\alpha} < s$, and so only the remainder term in \eqref{eqn:leading_term_sobolev_pf1} is non-zero; thus we have shown~\eqref{eqn:leading_term_sobolev} when $q = 1$ and $s = 1$.
+
+	When $s \geq 2$, we can analyze $E_{\alpha,x}$ using a Taylor expansion of $p$.  For any $x,z \in \Xset$, we have
 	\begin{equation*}
@@
 	\end{equation*}
-	and $p^{(\beta)} \in C^{\infty}(\Rd)$ additionally satisfies
-	\begin{equation*}
-	\norm{p^{(\beta)}}_{C^{s - \abs{\beta} - 1}(\Rd)} \leq p_{\max}.
-	\end{equation*}
-	Replacing $p$ by its Taylor expansion, the term $E_{\alpha,x}$ can be rewritten as
-	\begin{align*}
-	E_{\alpha,x} & = \int_{\Rd} (x - z)^{\alpha} K_r(x,z) p(z) \,dz \\
-	& =  \sum_{\abs{\beta} = 0}^{s - 2} \frac{p^{(\beta)}(x)}{\beta!} \int_{\Rd} (x - z)^{\alpha + \beta}K_r(x,z)\,dz ~~+ \\
-	& ~~~\frac{1}{(s - 2)!}\sum_{\abs{\beta} = s - 1} \int_{\Rd} \int_{0}^{1} (1 - t)^{s - 2} (x - z)^{\alpha + \beta - 1}  p^{(\beta)}(x + t(z - x)) K_r(x,z) \,dz \,dt \\
-	& = \sum_{\abs{\beta} = 0}^{s - 2} r^{\abs{\alpha} + \abs{\beta}} \frac{p^{(\beta)}(x)}{\beta!} \int_{\Rd} y^{\alpha + \beta}K(y)\,dy +~ \frac{r^{\abs{\alpha} + s - 1}}{(s - 2)!} \sum_{\abs{\beta} = s - 1} \int_{\Rd} \int_{0}^{1} (1 - t)^{s - 2} y^{\alpha + \beta - 1} p^{(\beta)}(x + ytr) K(\norm{y}) \,dy \,dt \\
-	& =: \sum_{\abs{\beta} = 0}^{s - 2} \1\{\abs{\alpha} + \abs{\beta} > 1\}r^{\abs{\alpha} + \abs{\beta}} p_{\abs{\beta}}(x) + r^s p_s(x) ~~\textrm{for $p_{\abs{\beta}} \in C^{\infty}(\Rd), \norm{p_{\abs{\beta}}}_{C^{s - \abs{\beta} - ,1}(\Rd)} \leq c \norm{p}_{C^{s - 1}(\Rd)}$},
-	\end{align*}
-	where the last line follows since $K$ is a 2nd order kernel. 
-	Substituting our expressions for $E_{\alpha,x}$ and $E_{\alpha,x,t}(f)$ back into~\eqref{eqn:leading_term_sobolev_pf1}, we obtain
-	\begin{align*}
-	\Ebb\bigl[D_kf(x)\big] & = \sum_{\abs{\alpha} = 1}^{s - 1} \sum_{\abs{\beta} = 0}^{s-2} \1\{\abs{\alpha} + \abs{\beta} > 1\} r^{\abs{\alpha} + \abs{\beta}} \frac{f^{(\alpha)}(x)p_{\abs{\beta}}(x)}{\alpha!} + r^s \biggl(\sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)p_s(x)}{\alpha!} + g_s(x)\biggr) \\
-	& = \sum_{\ell = 2}^{s - 1} r^{\ell} \sum_{ \substack{\abs{\alpha} + \abs{\beta} = \ell, \\ \abs{\beta} > 0} }   \frac{f^{(\alpha)}(x)p_{\abs{\beta}}(x)}{\alpha!} + r^s \biggl(\sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)p_s(x)}{\alpha!} + g_s(x)\biggr) \\
+	and $p^{(\beta)} \in C^{s - \abs{\beta} - 1}(\Xset)$ additionally satisfies
+	\begin{equation*}
+	\norm{p^{(\beta)}}_{C^{s - \abs{\beta} - 1}(\Xset)} \leq p_{\max}.
+	\end{equation*}
+	Replacing $p$ by its Taylor expansion, we analyze the term $E_{\alpha,x}$. Note that by assumption $\sigma > r$, and therefore for any $x \in \Xset_{\sigma}$,
+	\begin{equation*}
+	\int_{\Xset} (x - z)^{\alpha + \beta} K_r(x,z) \,dz = \int_{\Rd} (x - z)^{\alpha + \beta} K_r(x,z) \,dz = r^{\abs{\alpha} + \abs{\beta}} \int_{\Rd} y^{\alpha + \beta} K\bigl(\norm{y}\bigr) \,dz
+	\end{equation*} 
+	and similarly
+	\begin{equation*}
+	\int_{\Xset} \int_{0}^{1} (1 - t)^{s - 2} (x - z)^{\alpha + \beta}  p^{(\beta)}(x + t(z - x)) K_r(x,z) \,dz \,dt = r^{\abs{\alpha} + \abs{\beta}}\int_{\Rd} \int_{0}^{1} (1 - t)^{s - 2} y^{\alpha + \beta}  p^{(\beta)}(x + try) K\bigl(\norm{y}\bigr) \,dy \,dt
+	\end{equation*}
+	Therefore for any such $x \in \Xset_{\sigma}$,
+	\begin{align*}
+	E_{\alpha,x} & = \int_{\Xset} (x - z)^{\alpha} K_r(x,z) p(z) \,dz \\
+	& =  \sum_{\abs{\beta} = 0}^{s - 2} \frac{p^{(\beta)}(x)}{\beta!} \int_{\Xset} (x - z)^{\alpha + \beta}K_r(x,z)\,dz ~~+ \\
+	& ~~~\frac{1}{(s - 2)!}\sum_{\abs{\beta} = s - 1} \int_{\Xset} \int_{0}^{1} (1 - t)^{s - 2} (x - z)^{\alpha + \beta}  p^{(\beta)}(x + t(z - x)) K_r(x,z) \,dz \,dt \\
+	& = \sum_{\abs{\beta} = 0}^{s - 2} r^{\abs{\alpha} + \abs{\beta}} \frac{p^{(\beta)}(x)}{\beta!} \int_{\Rd} y^{\alpha + \beta}K(y)\,dy +~ \frac{r^{\abs{\alpha} + s - 1}}{(s - 2)!} \sum_{\abs{\beta} = s - 1} \int_{\Rd} \int_{0}^{1} (1 - t)^{s - 2} y^{\alpha + \beta} p^{(\beta)}(x + ytr) K\bigl(\norm{y}\bigr) \,dy \,dt \\
+	& =: \sum_{\abs{\beta} = 0}^{s - 2} \1\{\abs{\alpha} + \abs{\beta} > 1\}r^{\abs{\alpha} + \abs{\beta}} p_{\abs{\beta}}(x) + r^s p_{s - 1}(x) ~~\textrm{for $p_{\abs{\beta}} \in C^{\infty}(\Xset), \norm{p_{\abs{\beta}}}_{C^{s - \abs{\beta} - 1}(\Xset)} \leq c \norm{p}_{C^{s - 1}(\Xset)}$},
+	\end{align*}
+	where the last line follows since $K$ is a 2nd order kernel. On the other hand when $x \not\in \Xset_{\sigma}$, the derivatives $f^{(\alpha)}(x) = 0$ for each $\alpha = 1,2,\ldots,s-1$. 
+	As a result, by substituting our expressions for $E_{\alpha,x}$ and $E_{\alpha,x,t}(f)$ back into~\eqref{eqn:leading_term_sobolev_pf1}, we obtain that for all $x \in \Xset$,
+	\begin{align*}
+	\Ebb\bigl[D_kf(x)\big] & = \sum_{\abs{\alpha} = 1}^{s - 1} \sum_{\abs{\beta} = 0}^{s-2} \1\{\abs{\alpha} + \abs{\beta} > 1\} r^{\abs{\alpha} + \abs{\beta}} \frac{f^{(\alpha)}(x)p_{\abs{\beta}}(x)}{\alpha!} + r^s \biggl(\sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)p_{s - 1}(x)}{\alpha!} + g_s(x)\biggr) \\
+	& = \sum_{\ell = 2}^{s - 1} r^{\ell} \sum_{ \substack{\abs{\alpha} + \abs{\beta} = \ell, \\ \abs{\beta} > 0} }   \frac{f^{(\alpha)}(x)p_{\abs{\beta}}(x)}{\alpha!} + r^s \biggl(\sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)p_{s - 1}(x)}{\alpha!} + g_s(x)\biggr) \\
 	& =: \sum_{\ell = 2}^{s - 1} r^{\ell} f_{\ell}(x) + r^s f_s(x).
 	\end{align*}
-	It remains to show that $\norm{f_{\ell}}_{W^{s - \ell,2}(\Rd)} \leq c \norm{f}_{W^{s,2}(\Rd)}$ for each $\ell = 2,\ldots,s$.
-To see this, note that for $\abs{\alpha} + \abs{\beta} = \ell$ and $\min\{s - \abs{\alpha}, s - 1 - \abs{\beta}\} = m$, the inequality $s - \ell \leq m$ holds.
-Therefore,
+	We have already shown $g_s \in C^{0}(\Xset) \cap \Leb_{\sigma - r}^2(\Xset)$.
+Since additionally $f^{(\alpha)} \in C^{s - \abs{\alpha}}(\Xset) \cap W_{\sigma}^{s - \abs{\alpha},2}(\Xset)$ and $p_{\abs{\beta}} \in C^{s - 1 - \abs{\beta}}(\Xset)$, the products
+	\begin{equation*}
+	f^{(\alpha)} \cdot p_{\abs{\beta}} \in W_{\sigma}^{m,2}(\Xset) \cap C^{m}(\Xset) ~~\textrm{for $m = \min\{s - \abs{\alpha}, s - 1 - \abs{\beta}\}$.}
+	\end{equation*}
+	Finally, note that for $\abs{\alpha} + \abs{\beta} = \ell$ the inequality $s - \ell \leq m$ holds.
+Therefore, $f_{\ell} \in W_{\sigma}^{s - \ell,2}(\Xset) \cap C^{s - \ell}(\Xset)$ for each $\ell = 2,3,\ldots,s$, and
 	\begin{equation*}
@@
 
-	If $s \leq 2$, by hypothesis there exists $f_s$ satisfying
-	\begin{equation*}
-	\norm{f_s}_{\Leb^2(\Rd)} \leq c \norm{f}_{W^{s,2}(\Rd)}
-	\end{equation*}
-	such that for any $z \in \Rd$,
+	If $s \leq 2$, by hypothesis there exists $f_s \in C^{0}(\Xset) \cap \Leb_{\sigma - r}^{2}(\Xset)$ satisfying
+	\begin{equation*}
+	\norm{f_s}_{\Leb^2(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}
+	\end{equation*}
+	such that for any $z \in \Xset$,
 	\begin{equation*}
@@
 	\end{equation*}
-	and again using the inductive hypothesis, we have that
+	Seeing as $f_s \in C^{0}(\Xset) \cap \Leb_{\sigma - r}^2(\Xset)$, we may apply the inductive hypothesis to obtain that $\Ebb[D_kf_s(x)] \in C^0(\Xset) \cap \Leb_{\sigma - (q + 1)r}^2(\Xset)$, and additionally
 	\begin{equation}
 	\label{eqn:leading_term_sobolev_pf3}
-	\biggl\|\Ebb\Bigl[D_kf_s(x)\Bigr]\biggr\|_{\Leb^2(\Rd)} \leq c \norm{f_s}_{\Leb^2(\Rd)},
-	\end{equation}
-	establishing~\eqref{eqn:leading_term_sobolev} for all $q$ in the case when $s \leq 2$.
-
-	Otherwise, when $s \geq 3$ by hypothesis there exist functions $f_{\ell}$ for $\ell = 2,\ldots,s$ satisfying
-	\begin{equation*}
-	\norm{f_{\ell}}_{W^{s - \ell,2}(\Rd)} \leq c \norm{f}_{W^{s,2}(\Rd)}
-	\end{equation*}
-	such that for any $z \in \Rd$,
+	\biggl\|\Ebb\Bigl[D_kf_s(x)\Bigr]\biggr\|_{\Leb^2(\Xset)} \leq c \norm{f_s}_{\Leb^2(\Xset)}.
+	\end{equation}
+	We have therefore established~\eqref{eqn:leading_term_sobolev} for all $q$ in the case when $s \leq 2$.
+
+	Otherwise, when $s \geq 3$ by hypothesis there exist functions $f_{\ell} \in C^{s - \ell}(\Xset) \cap W_{\sigma - r}^{s - \ell,2}(\Xset)$ for each $\ell = 2,\ldots,s$ satisfying
+	\begin{equation*}
+	\norm{f_{\ell}}_{W^{s - \ell,2}(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}
+	\end{equation*}
+	such that for any $z \in \Xset$,
 	\begin{equation*}
@@
 	\end{equation*}
-	As a result, by the law of iterated expectation
+	By the law of iterated expectation
 	\begin{align*}
@@
 	First we consider the case when $2(q + 1) \geq s$.
-Note that $\ell \geq 2$ implies $2q \geq s - \ell$.
-Therefore by hypothesis, for each $\ell = 2,\ldots,s - 1$ the expectation $\Ebb[D_kf_{\ell}(x)] = r^{s - \ell}f_{\ell,s}(x)$ for some $f_{\ell,s} \in \Leb^2(\Rd)$ satisfying
+Note that $2q \geq s - \ell$ for each $\ell \geq 2$.
+Therefore by hypothesis, for each $\ell = 2,\ldots,s - 1$ the expectation $\Ebb[D_kf_{\ell}(x)] = r^{s - \ell}f_{\ell,s}(x)$ for some $f_{\ell,s} \in C^{s - \ell}(\Xset) \cap \Leb_{\sigma - (q + 1)r}^2(\Rd)$ satisfying
 	\begin{equation}
 	\label{eqn:leading_term_sobolev_pf4}
-	\norm{f_{\ell,s}}_{\Leb^2(\Rd)} \leq c\norm{f_{\ell}}_{W^{s - \ell,2}(\Rd)} \leq c\norm{f}_{W^{s,2}(\Rd)}
+	\norm{f_{\ell,s}}_{\Leb^2(\Xset)} \leq c\norm{f_{\ell}}_{W^{s - \ell,2}(\Xset)} \leq c\norm{f}_{W^{s,2}(\Xset)}
 	\end{equation}
@@
 
-	Otherwise $2(q + 1) < s - 1$.
-For each $\ell = 2,\ldots, s - 1$, if additionally  $2q \leq s - \ell - 1$, then by hypothesis $\Ebb(D_kf_{\ell}(x)) = \sum_{m = 2q}^{s - \ell - 1} r^{m} \cdot f_{\ell, \ell + m}(x) + r^{s - {\ell}} f_{\ell,s}$, and otherwise $\Ebb(D_kf_{\ell}(x)) = r^{s - \ell}f_{\ell,s}(x)$, where for each $\ell$ and $m$,
-	\begin{equation*}
-	\norm{f_{\ell,\ell + m}}_{W^{s - (\ell + m),2}(\Rd)} \leq c\norm{f_{\ell}}_{W^{s - \ell}} \leq c\norm{f}_{W^{s,2}(\Rd)}
-	\end{equation*}
-	and $f_{\ell,s}$ satisfies \eqref{eqn:leading_term_sobolev_pf4}. Therefore,
+	Otherwise $2(q + 1) < s - 1$. For each $\ell = 2,\ldots, s - 1$, if additionally  $2q \leq s - \ell - 1$, then by hypothesis
+	\begin{equation*}
+	\Ebb(D_kf_{\ell}(x)) = \sum_{m = 2q}^{s - \ell - 1} r^{m} \cdot f_{\ell, \ell + m}(x) + r^{s - {\ell}} f_{\ell,s}
+	\end{equation*}
+	where $f_{\ell,\ell + m} \in C^{s - (\ell + m)}(\Xset) \cap W_{\sigma - (q + 1)r}^{s - (\ell + m),2}(\Xset)$ satisfies
+	\begin{equation*}
+	\norm{f_{\ell,\ell + m}}_{W^{s - (\ell + m),2}(\Rd)} \leq c\norm{f_{\ell}}_{W^{s - \ell}} \leq c\norm{f}_{W^{s,2}(\Rd)}.
+	\end{equation*}
+	On the other hand if $2s > s - \ell - 1$, then
+	\begin{equation*}
+	\Ebb\bigl[D_kf_{\ell}(x)\bigr] = r^s f_{\ell,s}
+	\end{equation*}
+	for some $f_{\ell,s} \in C^0(\Xset) \cap \Leb_{\sigma - (q + 1)r}^{2}(\Xset)$ which additionally satisfies \eqref{eqn:leading_term_sobolev_pf4}. Therefore,
 	\begin{align*}
@@
 \end{proof}
+
+\subsection{One-Sided Concentration}
 

### 2020-01-30 notes/weekly_notes/1.30.20.MeetingNotes.tex
added higher-order testing rate overly compactly supported Sobolev functions

--- 2020-01-30 notes/graph_testing.tex
@@
 \begin{document}
-\title{Graph Testing}
+\title{Meeting Notes for Week 1/23/20 - 1/30/20}
 \author{Alden Green}
@@
 
-This document details the current status of the graph testing project.
-We divide by section based on the testing problem under consideration.
-At the end of each section, we list some areas we are interested in investigating.
-All proofs are left until the end.
-
-We establish some notation that we will use throughout.
-A graph $G = (V,W)$ consists of a collection of vertices $V = \{v_1,\ldots,v_n\}$, and a weight matrix $W \in \Reals^{\abs{V} \times \abs{V}}$ with $ij$th entry $W_{ij}$ encoding affinity between $v_i$ and $v_j$.
-We will sometimes write $W_G$ to make it clear which graph a weight matrix is associated with.
-When the weight matrix $W_G$ consists only of $0$s and $1$s, we will use the equivalent representation $G = (V,E(G))$ when convenient; here $E(G) = V \times V \cap \{(v_i,v_j): W_{ij} = 1\}$.
-
-We will focus our attention on neighborhood graphs, and operators associated with them.
-For a kernel $K:\Reals \to \Reals$ and radius $r > 0$, we define $G_{n,K}=(V,W)$ to be the neighborhood graph with vertices $V = X$, and weights $W_{ij} = K(\norm{x_i - x_j}/r)$.
-Of particular interest to us will be the \emph{random geometric graph} $G_{n,r}$, the neighborhood graph constructed using the uniform kernel $K(z) := \1\{z \leq 1\}$.
-The weight matrix $W(G_{n,r})$ consists only of $0$s and $1$s, and we will thus use the equivalent representation $G_{n,r} = (X,E(G_{n,r}))$.
-
-The primary operator we are interested in (at this point) is the graph Laplacian. Let $D_G$ be the diagonal degree
-matrix of a graph $G$, with entries $D_{uu} := \sum_{v \in V} W_{uv}$.
-The graph Laplacian is then $L_G = D_G - W_G$.
-Let $\Lambda(G)$ consist of the eigenvalues $0 = \lambda_1(G) \leq \lambda_2(G) \leq \cdots \leq \lambda_n(G)$ of $L_G$, with $v_k(G) = (v_{k,1}(G),\ldots,v_{k,n}(G)) \in \Reals^n$ denoting the eigenvector corresponding to the $k$th eigenvalue $\lambda_k(G)$.
-
-\section{Regression goodness-of-fit testing.}
-
-Suppose we observe samples $(y_i,x_i)$ for $i = 1,\ldots,n$, where conditional on $X$ the responses $Y = \{y_1,\ldots,y_n\}$ are assumed to follow the model
-\begin{equation}
-\label{eqn:regression_known_variance}
+Suppose we observe samples $(y_i,x_i)$ for $i = 1,\ldots,n$, where $x_1,\ldots,x_n$ are sampled independently from a distribution $P$ with density $p$, and conditional on $X$ the responses $Y = \{y_1,\ldots,y_n\}$ are assumed to follow the model
+\begin{equation}
+\label{eqn:regression_random_design_known_variance}
 y_i = f(x_i) + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
 \end{equation} 
-and our task is to distinguish
+Our task is to distinguish
 \begin{equation*}
@@
 \end{equation*}
-
-We will evaluate our performance using worst-case risk: for a given function class $\mathcal{H}$ and test function $\phi: \Reals^n \to \set{0,1}$, let
-\begin{equation*}
-\mathcal{R}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, f \neq f_0} \Ebb_f(1 - \phi).
-\end{equation*}
-The worst-case risk may be quite close to $1$ unless we enforce some separation between null and alternative spaces. A more realistic measure of performance is therefore
+and we worst-case risk to assess performance: for a test $\phi$ and function class $\mathcal{H}$,
 \begin{equation*}
@@
 
-\subsection{Test Statistics.}
-We list the test statistics we use for the regression testing problem.
-
-\paragraph{Eigenvector projection test statistic.}
-The graph Laplacian eigenvector projection test is a truncated-series test.
-Suppose we are given a graph $G = (X,W)$ defined on the sample points $X$.
-Letting $\kappa$ be some integer between $1$ and $n$, our test projects $Y$ onto the eigenvectors $v_1(G),\ldots,v_{\kappa}(G)$, and then takes the empirical norm of this projection to be the test statistic.
-Formally,
-\begin{equation*}
-T_{\mathrm{spec}}(G) := \frac{1}{n} \sum_{k = 1}^{\kappa} \Biggl(\sum_{i = 1}^{n} y_i v_{k,i}(G)\Biggr)^2
-\end{equation*}
-Then the test $\phi_{\spec}(G) := \1\{T_{\spec}(G) \geq \tau\}$ rejects the null hypothesis when the test statistic $T_{\mathrm{spec}}(G)$ is greater than a pre-specified cutoff $\tau$.
-
-\subsection{Random design.}
-In the random design case, we assume that each design point $x_i$ is sampled independently from some distribution $P$. Formally, our model for the samples is the following:
-\begin{equation}
-\label{eqn:regression_random_design_known_variance}
-x_1,\ldots,x_n \overset{\textrm{i.i.d}}{\sim} P,~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)~\textrm{and}~ \varepsilon \perp X,~~ y_i = f(x_i) + \varepsilon_i.
-\end{equation}
-
-In Theorem~\ref{thm:sobolev_testing_rate_order1} we show that under some typical regularity conditions on $P$, the neighborhood graph Laplacian eigenvector projection test $\phi_{\textrm{spec}}(G_{n,r})$ is a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};R)$ for $d = 1,2,3$ and $\Xset = [0,1]^d$.
-Let $p_d$ equal $3/4$ when $d = 2$ and equal $1/d$ for $d \geq 3$.
-
-\begin{theorem}
-	\label{thm:sobolev_testing_rate_order1}
-	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}.
-Let $R > 0$ and $b \geq 1$ be fixed constants, and $d = 1,2$ or $3$.
-Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p(x)$ is bounded away from zero and infinity,
-	\begin{equation*}
-	0 < p_{\min} < p(x) < p_{\max} < \infty,~~ \textrm{for all $x \in \mathcal{X}$.}
-	\end{equation*}
-	and the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choices
-	\begin{equation*}
-	c \frac{(\log n)^{p_d}}{n^{1/d}} \leq r(n) \leq n^{-4/((4 + d)(2+d))}, ~\kappa = (nR)^{2d/(4 + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
-	\end{equation*}
-	for $c$ a constant that depends only on $\Xset, p_{\min}$ and $p_{\max}$.
-Then the following statements holds for every $n$ sufficiently large: there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
-	\begin{equation}
-	\label{eqn:sobolev_testing_rate_order1}
-	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4 + d)}\} \cdot n^{-4/(4 + d)}
-	\end{equation}
-	the worst-case risk is upper bounded
-	\begin{equation}
-	\label{eqn:sobolev_testing_rate_order1_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}(G_{n,r}); W^{1,2}(\mathcal{X};R)) \leq \frac{c_2}{b}.
-	\end{equation}
-\end{theorem}
-
-The testing problem in the random design case with known variance fundamentally changes in the low-smoothness regime, where the number of bounded derivatives is small compared to the dimension $d$.
-In the case where $f \in W^{1,2}(\Xset;R)$ is assumed to have one derivative which is bounded in $\Leb^2(\Xset)$ norm, this transition occurs at $d = 4$.
-We shall discuss testing in the low-smoothness regime shortly.
-
-Theorem~\ref{thm:sobolev_testing_rate_order1} establishes that a truncated series test using graph eigenvectors performs just as well, in a minimax sense, as a truncated series test using a classic Fourier series; for analysis of the latter test as well as derivation of the minimax rate of nonparametric goodness-of-fit testing over Sobolev classes see \textcolor{red}{(Ingster)}.
-Intuitively, we can understand this by thinking of the graph eigenvectors as estimates of eigenfunctions of a density-weighted continuum Laplacian operator.
-When the density is uniform, and the domain $\Xset$ is the unit cube, under appropriate boundary conditions these eigenfunctions are themselves elements of the Fourier basis.
-When the density is merely close to uniform, the eigenfunctions no longer belong to the Fourier basis; nevertheless, they share enough properties with the latter that the resulting minimax rate is unchanged.
-Theorem~\ref{thm:sobolev_testing_rate_order1} shows that the additional error due to using \textit{estimates} of these eigenfunctions does not fundamentally change the minimax rate.
-
-When we additionally believe $W^{s,2}(\Xset)$, we know (again from Ingster) that the minimax testing rate is $n^{-4s/(4s + d)}$.
-Now, however, our estimates of the eigenvectors no longer track the derivatives finely enough at the boundaries of $\Xset$ to establish the same minimax rate over all $W^{s,2}(\mathcal{X})$.
-However, if we assume $f$ and its derivatives are supported only on the interior of $\mathcal{X}$, we recover the usual rate.
-Theorem~\ref{thm:sobolev_testing_rate} presents our formal result, that $\phi_{\textrm{spec}}(G_{n,r})$ is a minimax optimal test over the Sobolev balls $W_{\sigma}^{s,2}(\mathcal{X};R)$ whenever $4s > d$ and $\Xset = [0,1]^d$.
+Theorem~\ref{thm:sobolev_testing_rate} presents our formal result, that the graph spectral test 
+\begin{equation*}
+\phi_{\spec}(G_{n,r}) := \1\bigl\{T_{\textrm{spec}}(G_{n,r}) \geq \tau\bigr\},~~ T_{\textrm{spec}}(G_{n,r}) := \frac{1}{n}\sum_{k = 1}^{\kappa} \biggl(\sum_{i = 1}^{n} v_{k,i}(G_{n,r}) y_i\biggr)
+\end{equation*} is a minimax optimal test over the Sobolev balls $W_{\sigma}^{s,2}(\mathcal{X};R)$ whenever $4s > d$ and $\Xset = [0,1]^d$.
+
+Here, $W_{\sigma}^{s,2}(\Xset;L)$ consists of all those functions $f \in W^{s,2}(\Xset)$ such that $\mathrm{supp}(f^{(\alpha)}) \subseteq \Xset_{\sigma}$ for each $\abs{\alpha} \leq s$.
 
@@
 	\end{equation*}
-	Then the following statement holds for all $n$ sufficiently large: if the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choice
+	Then the following statement holds for all $n$ sufficiently large: if the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choices
 	\begin{equation*}
@@
 
-The restriction that $p$ be supported on the unit cube is mostly for convenience.
-If instead $p$ is supported on any compact set with Lipschitz boundary, the Theorem statement should hold (up to constants), with only a few modifications to our proofs.
-However, we do not work through the details.
-
-The restriction $4s > d$ is a fundamental consequence of the tightness of the Sobolev embedding theorem. When $4s \geq d$ the compact embedding
-\begin{equation*}
-W_d^{s,2}(\mathcal{X}) \subseteq \mathcal{L}_d^4(\mathcal{X}) 
-\end{equation*}
-does not hold (as it does when $d < 4s$). However, if we directly assume $f \in \mathcal{L}_d^4(\mathcal{X};R)$ (regardless of $d$) we obtain the following result.
-\begin{proposition}
-	\label{prop:L4_testing_rate}
-	If $f \in \Leb^4(\mathcal{X};R)$, there exists a constant $c$ such that if
-	\begin{equation}
-	\label{eqn:L4_testing_rate}
-	\epsilon^2 > b \cdot n^{-1/2}
-	\end{equation}
-	then the test
-	\begin{equation*}
-	\phi_{\mathrm{mean}} = \1\{\frac{1}{n}\sum_{i = 1}^{n} y_i^2 \geq 1\}
-	\end{equation*}
-	has worst-case risk
-	\begin{equation*}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}. 
-	\end{equation*}
-\end{proposition}
-Note that when $4s > d$ the critical radius~\eqref{eqn:sobolev_testing_rate} is smaller than \eqref{eqn:L4_testing_rate}.
-
-\subsection{Analysis}
-
-To prove Theorem~\ref{thm:sobolev_testing_rate}, we show that there exists a high-probability set $E \subseteq \Xset^n$ such that conditional on $X \in E$, the test $\phi_{\spec}(G_{n,r})$ has small risk.
-Since $G_{n,r}$ is a function only of $X$ and not of $Y$, this amounts to reasoning about the behavior of the test $\phi_{\spec}$ over a fixed graph $G = (X,E)$, where we observe
-\begin{equation}
-y_i = \beta_i + \varepsilon_i,~~\varepsilon_i \sim \mathcal{N}(0,1)
-\end{equation}
-for some fixed $\beta \in \Reals^n$.
-In Lemma~\ref{lem:fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}(G)$.
-Our bound on the Type II error will be stated as a function of $\beta^T L^s \beta$--a measure of the smoothness the signal $\beta$ displays over the graph $G$--as well as the $\kappa$th eigenvalue $\lambda_{\kappa}$.
-
-\begin{lemma}
-	\label{lem:fixed_graph_testing}
-	Let $1 \leq \kappa \leq n$ be an integer.
-	\begin{enumerate}
-		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}$ is upper bounded
-		\begin{equation}
-		\label{eqn:graph_spectral_type_I_error}
-		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{1}{b^2}.
-		\end{equation}
-		\item \textbf{Type II error:} For any $b \geq 1$ and $\beta$ such that
-		\begin{equation}
-		\label{eqn:fixed_graph_testing_critical_radius}
-		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{\beta^T L^s \beta}{n\lambda_{\kappa}^s}
-		\end{equation}
-		the Type II error of $\phi_{\spec}$ is upper bounded,
-		\begin{equation}
-		\label{eqn:graph_spectral_type_II_error}
-		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{3}{b}.
-		\end{equation}
-	\end{enumerate}
-\end{lemma}
-
-\subsection{Areas to Investigate}
-
-\begin{enumerate}[(i)]
-	\item When $d \geq 4$, is it true that there does not exist a uniformly consistent
-	test over the Sobolev ball $W_d^{1,2}(\mathcal{X};R)$?
-	\item Assume the distribution $P$ is supported on a manifold $\mathcal{M}$ of intrinsic dimension $s < d$.
-Does $\mathrm{\phi_{\mathrm{spec}}}$ display adaptivity to the intrinsic dimension of $\mathcal{M}$?
-	\item Assume that $f$ belongs to the Holder space $C_d^s(\mathcal{X})$.
-Moreover, suppose that instead of observing ${y_i}$ according to the regression testing model \eqref{eqn:regression_known_variance}, we observe
-	\begin{equation*}
-	y_i = f(x_i) + \sigma \varepsilon_i, ~\varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
-	\end{equation*}
-	where $\sigma > 0$ is unknown.
-When $d \geq 4$, what are the minimax regression testing rates over $C_d^1(\mathcal{X};R)$?
-Is the test $\phi_{\mathrm{spec}}$ minimax optimal, when the tuning parameters $r$ and $\kappa$ are appropriately chosen?
-\end{enumerate}
-
-\section{Two-sample density testing.}
-In the two-sample density testing problem, we observe independent samples $Z = z_1,\ldots,z_N \sim P$ and $Y = y_1,\ldots,y_M \sim Q$, where $P$ and $Q$ are distributions over $\Reals^d$ with densities $p$ and $q$, respectively, and $N \sim \textrm{Bin}(n,1/2)$.
-Our goal is to distinguish the hypotheses
-\begin{equation*}
-\mathbf{H}_0: P = Q \quad \textrm{vs} \quad \mathbf{H}_{\textrm{a}}: P \neq Q
-\end{equation*}
-and we again evaluate our performance using worst-case risk; letting $\phi:\Reals^{N + M} \to \{0,1\}$, 
-\begin{equation*}
-\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \inf_{p \in \mathcal{H}}\Ebb_{p,p}(\phi) + \sup_{\substack{p,q \in \mathcal{H} \\ \norm{p - q}_{\Leb^2} \geq \epsilon}} \Ebb_{p,q}(1 - \phi).
-\end{equation*}
-
-\subsection{Test statistics.}
-
-We suggest several two-sample test statistics.
-
-\paragraph{Eigenvector projection test statistic.}
-
-It is straightforward to adapt the test statistic $T_{\mathrm{spec}}$ to the two-sample testing problem.
-Concatenate the samples $Z$ and $Y$ in $X = (z_1,\ldots,z_N,y_1,\ldots,y_M)$, and define $T_{\mathrm{spec}}^{(2)}$ to be
-\begin{equation}
-\label{eqn:graph_spectral_projections_2}
-T_{\mathrm{spec}}^{(2)} := \frac{1}{n} \sum_{k = 0}^{\kappa} \left(\sum_{i = 1}^{n} v_i a_i\right)^2, ~~\textrm{where}~~ a = (\underbrace{N^{-1},\ldots,{N^{-1}}}_{\textrm{length } N},\underbrace{-M^{-1},\ldots,-M^{-1}}_{\textrm{length } M})
-\end{equation}
-
-For convenience, we state our following two test statistics with respect to the empirical norm $\norm{\theta}_n = n^{-1/2}\norm{\theta}_2$ for $\theta \in \Reals^n$.
-They will each depend on a tuning parameter $\lambda > 0$.
-\paragraph{Graph Sobolev IPM.}
-Letting $C_n := nr^{(d + 2)/2}$ and
-\begin{equation*}
-\Theta_{1,2} := \{\theta \in \Reals^n:~ C_n^{-1} \norm{B\theta}_2 \leq 1\} 
-\end{equation*}
-we define the \emph{graph Sobolev IPM} to be
-\begin{equation}
-\label{eqn:sobolev_IPM}
-T_{\textrm{sob}} := \sup_{\substack{\theta \in \Theta_{1,2} \\ \lambda \norm{\theta}_n \leq 1}} \abs{\frac{1}{n}\sum_{i = 1}^{n} a_i \theta_i}. 
-\end{equation}
-
-\paragraph{Graph Total Variation IPM.}
-Letting $C_n' := n^{2}r^{(d + 1)}$ and 
-\begin{equation*}
-\Theta_{1,1} := \{\theta \in \Reals^n:~ (C_n')^{-1} \norm{B\theta}_1 \leq 1\}
-\end{equation*}
-we define the \emph{graph Total Variation} IPM to be
-\begin{equation}
-\label{eqn:total_variation_IPM}
-T_{\mathrm{TV}} := \sup_{\substack{\theta \in \Theta_{1,1}, \\ \lambda \norm{\theta}_n \leq 1} } \abs{\frac{1}{n}\sum_{i = 1}^{n} a_i \theta_i}, \quad
-\end{equation}
-
-\subsection{Current results.}
-
-In Theorem~\ref{thm:twosample_sobolev_testing_rate} we show that under some typical regularity conditions on $P$, the test $\phi_{\textrm{spec}}^{(2)} := \1\{T_{\mathrm{spec}}^{(2)} \geq \tau\}$ is, up to log factors, a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};R)$ when $d = 1$.
-
-\begin{theorem}
-	\label{thm:twosample_sobolev_testing_rate}
-	Let $b \geq 1$ and $a > 0$ be fixed constants, and let $d = 1$.
-Suppose that $\mu = (P + Q)/2$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]$ with density functions $\rho$ bounded above and below by constants, i.e
-	\begin{equation*}
-	0 < \rho_{\min} < \rho(x) < \rho_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
-	\end{equation*}
-	Then the following statement holds: if the test $\phi_{\spec}^{(2)}$ is performed with parameter choices 
-	\begin{equation*}
-	r = \log^a n \cdot \left(\frac{\log n}{n}\right), ~\kappa = n^{2/5}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
-	\end{equation*}
-	then there exists a constant $c$ which may depend on $R,p_{\max},q_{\max},b$ and $a$ but is independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
-	\begin{equation}
-	\label{eqn:twosample_sobolev_testing_rate}
-	\epsilon^2 \geq c \cdot b \cdot n^{-4/5} (\log n)^{h(a,1)}
-	\end{equation}
-	the worst-case risk is upper bounded
-	\begin{equation}
-	\label{eqn:twosample_sobolev_testing_rate_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; \mathcal{W}^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}.
-	\end{equation}
-\end{theorem}
-
-We prove Theorem~\ref{thm:twosample_sobolev_testing_rate} by relating the density testing problem to a regression testing problem with a certain type of structured noise, and then proceeding along similar lines to the proof of Theorem~\ref{thm:sobolev_testing_rate}.
-To pursue this strategy, we require the eigenvectors to satisfy a certain type of incoherence condition; this is in constrast to the regression testing problem with known variance, where we did not require the eigenvectors to be smooth in any sense.
-
-\subsection{Areas to Investigate}
-
-\begin{enumerate}[(i)]
-	\item When $1 < d < 4$, is the test $\phi_{\spec}^{(2)}$ minimax optimal?
-	\item Are either or both of the test statistics \eqref{eqn:graph_spectral_projections_2}-\eqref{eqn:sobolev_IPM} minimax optimal over $C_d^1(\mathcal{X};R)$ for all values of $d$?
-	\item Are either or both of the test statistics \eqref{eqn:graph_spectral_projections_2}-\eqref{eqn:sobolev_IPM} minimax optimal over $W_d^{1,2}(\mathcal{X};R)$ when $d \leq 4$?
-	\item Is the test statistic \eqref{eqn:graph_spectral_projections_2}, computed over a graph with suitable choice of kernel $K$, minimax optimal over higher order derivative classes $C_d^s$ and $W^{s,2}$?
-	\item Modify the test statistic \eqref{eqn:sobolev_IPM} by replacing the function class $\Theta_{1,2}$ with
-	\begin{equation}
-	\Theta_{s,2} := \{\theta \in \Reals^n:~ C_n^{-1} \norm{B^{(s)}\theta}_2 \leq 1\} 
-	\end{equation}
-	Is the modified test statistic, computed over a graph with suitable choice of kernel $K$, minimax optimal over higher order derivative classes $C_d^s$ and $W^{s,2}$?
-	\item What is the minimax testing rate over $BV_d^{1}(\mathcal{X};R)$? Does it exhibit a phase transition analogous to the minimax estimation rate over bounded variation spaces?
-	\item Is the test statistic \eqref{eqn:total_variation_IPM} minimax optimal over $BV_d^{1}(\mathcal{X};R)$?
-	\item Is the test statistic \eqref{eqn:total_variation_IPM} minimax optimal over Sobolev and Holder function classes?
-\end{enumerate}
-
-\section{Definitions and Notation}
-
-Here we collect definitions of some common function spaces and graph operators.
-
-\subsection{Function Spaces}
-
-\paragraph{Lebesgue spaces.}
-
-We say a Borel measurable function $f: \mathcal{X} \to \Reals$ is in the space $\mathcal{L}^p(\mathcal{X})$ for $1 \leq p < \infty$ if 
-$$\norm{f}_{\mathcal{L}^p(\mathcal{X})} := \int_{\mathcal{X}} \abs{f(x)}^p \,dx < \infty$$
-and we let 
-\begin{equation*}
-\mathcal{L}^p(\mathcal{X};R) = \set{f \in \mathcal{L}^p(\mathcal{X}): \norm{f}_{\mathcal{L}^p(\mathcal{X})} < R}
-\end{equation*}
-be a ball in the Lebesgue space.
-
-\paragraph{Holder spaces.}
-
-For a given $s > 0$, let $\ell = \floor{s}$ be the largest integer strictly less than $s$. Then the $s$th Holder norm is given by
-\begin{equation*}
-\norm{f}_{C_d^{s}(\mathcal{X})} := \sum_{\abs{\alpha} < s} \norm{D^{\alpha}f}_{\infty} + \sum_{\abs{\alpha} = \ell} \sup_{x,y \in \mathcal{X}} \frac{\abs{D^{\alpha}f(y) - D^{\alpha}f(x)}}{\norm{x - y}_2^{s - \ell}}
-\end{equation*}
-and the $s$th Holder space $C_d^{s}(\mathcal{X})$ consists of all functions which are $\ell$ times continuously differentiable with finite Holder norm $\norm{\cdot}_{C_d^{s}(\mathcal{X})}$.
-Denote the Holder ball by $C_d^{s}(\mathcal{X},R) = \set{f \in C_d^{s}(\mathcal{X}): \norm{f}_{C_d^{s}(\mathcal{X})} \leq R}$.
-
-\paragraph{Sobolev spaces.}
-
-For a given $s > 0$, the Sobolev space $W^{s,2}(\mathcal{X})$ consists of all functions $f \in \Leb^2(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^2(\mathcal{X})$.
-The Sobolev $\{s,2\}$ norm is then
-\begin{equation*}
-\norm{f}_{W^{s,2}(\mathcal{X})}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^2 \,dx
-\end{equation*}
-and for a given $R > 0$, the corresponding ball is $W^{s,2}(\Xset; R) = \set{f: \norm{f}_{W^{s,2}(\Xset)} \leq R}$.
-
-For a given $\sigma > 0$, let $\Xset_{\sigma} := \set{x \in \Xset: \mathrm{dist}(x,\partial\Xset) > \sigma}$.
-Let $W_{\sigma}^{s,2}(\Xset;L)$ consist of all those functions $f \in W^{s,2}(\Xset)$ such that $\mathrm{supp}(f) \subseteq \Xset_{\sigma}$.
-
-\paragraph{Bounded Variation spaces.}
-
-For a function $f \in L^1(\mathcal{X})$ the \emph{total variation} semi-norm of $f$ is
-\begin{equation*}
-TV(f;\mathcal{X}) := \sup \left\{ \int_{\mathcal{X}} f \, \Xsetive \, \psi \,dx : \psi \in C_c^1(\mathcal{X}; \Reals^d), \abs{\psi} \leq 1 \right\};
-\end{equation*}
-and we write $BV_d(\mathcal{X})$ for the subset of functions $f \in L^1(\mathcal{X})$ which have bounded norm
-\begin{equation*}
-\norm{f}_{BV_d(\mathcal{X})} := \norm{f}_{\infty} + TV(f;\mathcal{X}).
-\end{equation*}
-For a given $R > 0$, the corresponding ball is $BV_d^{1}(\mathcal{X};R) = \set{f: \norm{f}_{BV_d(\mathcal{X})} \leq R}$.
-
-\subsection{Graph Operators}
-Let $s \geq 1$ be an integer. The $s$th-order difference operator on $G_{n,r}$, denoted $B^{(s)}$, is defined by
-\begin{equation*}
-B^{(s)} :=
-\begin{cases}
-L^{s/2},& ~~ s \textrm{ even} \\
-BL^{(s - 1)/2},& ~~ s \textrm{ odd.}
-\end{cases}
-\end{equation*}
-
-\subsection{Kernels}
-We say a kernel function $K(\cdot)$ is a $2$nd order kernel if $K$ is compactly supported on $B(0,1)$, uniformly upper bound $\abs{K(x)} \leq K_{\max} < \infty$ for all $x \in B(0,1)$, and
-\begin{equation*}
-\int K(x) \,dx = \nu_d,~~ \int x K(x) \,dx = 0.
-\end{equation*}
-Note that the uniform kernel, defined as $K(x) = \1\{\norm{x} \leq 1\}$, is a $2$nd order kernel.
-
-\subsection{Notation}
-
-We will treat $f$ interchangeably as a function $f:\Rd \to \Reals$, and as a vector $f = (f(x_1),\ldots,f(x_n))$. It should always be clear from context how we are using $f$.
-For sequences $(a_n), (b_n)$, we say $a_n = O(b_n)$ if there exists a constant $c$ such that $a_n \leq cb_n$ for all $c$. We use the notation $a \wedge b$ for the minimum of $a$ and $b$.
-
-We use the notation $\Lambda(H)$ to denote the spectrum of a matrix $H$, and $\lambda_k(H)$ to denote the $k$th smallest eigenvalue of $H$.
-
 \section{Proofs}
 
-\subsection{Proof of Lemma~\ref{lem:fixed_graph_testing}}
-
-In this proof, we will drop all notational dependence on the graph $G$ for ease of reading.
-
-To prove Lemma~\ref{lem:fixed_graph_testing} we will first compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}$, and then use Chebyshev's inequality to show \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.
-
-\paragraph{Mean of $T_{\mathrm{spec}}$:} Using the notation $\dotp{v}{w} = \sum_{i = 1}^{n} v_iw_i$, we have
-\begin{align*}
-\Ebb(T_{\spec}) & = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{\beta}{v_k}^2 + \Ebb\bigl( \dotp{\varepsilon}{v_k}^2 + 2 \dotp{\varepsilon}{v_k} \dotp{\beta}{v_k}\bigr)\right) \\
-& = \frac{\kappa}{n} + \frac{1}{n}\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2.
-\end{align*}
-When $\beta = 0$, this equals $\kappa/n$. Otherwise, we have the following lower bound:
-\begin{align*}
-\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2 & = \norm{\beta}_2^2 - \sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \\
-& \geq \norm{\beta}_2^2 - \frac{1}{\lambda_{\kappa}^s}\sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \lambda_k^s \\
-& \geq \norm{\beta}_2^2 - \frac{\beta^T L^s \beta}{\lambda_{\kappa}^s},
-\end{align*}
-and therefore $\Ebb(T_{\spec}) \geq \kappa/n + n^{-1}(\norm{\beta}_2^2 - \beta^T L^s \beta/\lambda_{\kappa}^s)$.
-
-\paragraph{Variance of $T_{\mathrm{spec}}$:}
-We write $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvector $v_k$ as its $k$th column.
-\begin{align*}
-\Var(T_{\spec}) & = \frac{1}{n^2} \Var(y^T V_{\kappa} V_{\kappa}^T y) \\
-& = \frac{1}{n^2} \Var((\beta + \varepsilon)^T V_{\kappa} V_{\kappa}^T (\beta + \varepsilon)) \\
-& = \frac{1}{n^2} \Var(2 \beta^T V_{\kappa} V_{\kappa}^T \varepsilon + \varepsilon^T V_{\kappa} V_{\kappa}^T \varepsilon) \\
-& \leq \frac{1}{n^2}(4 \beta^T V_{\kappa} V_{\kappa}^T \beta + 2\kappa)
-\end{align*}
-where the last inequality follows from standard properties of the Gaussian distribution.
-We now move on to showing the desired inequalities \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.
-
-\paragraph{Proof of~\eqref{eqn:graph_spectral_type_I_error}:} By Chebyshev's inequality,
-\begin{align*}
-\Pbb_{\beta = 0}\bigl(T_{\spec} \geq \frac{\kappa}{n} + t(b)\bigr)
-& \leq \Pbb_{\beta = 0}\bigl(\abs{T_{\spec} - \frac{\kappa}{n}} \geq t(b)\bigr) \\
-& \leq \frac{\Var_{\beta = 0}(T_{\spec})}{t(b)^2} = \frac{1}{b^2}.
-\end{align*}
-
-\paragraph{Proof of~\eqref{eqn:graph_spectral_type_II_error}:} For simplicity, we introduce the notation
-\begin{equation*}
-\Delta = \frac{\norm{\beta}_2^2}{n} - \frac{\beta^T L^s \beta}{n\lambda_{\kappa}^s}.
-\end{equation*}
-Assumption~\eqref{eqn:fixed_graph_testing_critical_radius} implies $\Delta \geq 2 t(b)$. Then another application of Chebyshev's inequality gives us
-\begin{align*}
-\Pbb_{\beta}\bigl(T_{\spec} \leq \frac{\kappa}{n} + t(b)\bigr) & = \Pbb_{\beta}\bigl(T_{\spec} - \Ebb_{\beta}(T_{\spec}) \leq t(b) - \Delta \bigr) \\
-& \leq \Pbb_{\beta}\bigl(\abs{T_{\spec} - \Ebb_{\beta}(T_{\spec})} \leq \Delta - t(b) \bigr) \tag{since $\Delta \geq t(b)$}	\\
-& \leq \frac{\Var_{\beta}(T_{\spec})}{(\Delta - t(b))^2} \\
-& \leq 4\frac{\Var_{\beta}(T_{\spec})}{\Delta^2} \tag{since $\Delta \geq 2t(b)$} \\
-& \leq 4\frac{2\kappa/n^2 + 4\beta^T V_{\kappa} V_{\kappa}^T \beta /n^2}{\Delta^2}.
-\end{align*}
-We handle each summand in the numerator separately. For the first term, since $\Delta \geq 2 t(b)$, we have
-\begin{equation}
-\label{eqn:spectral_type_II_error_pf1}
-\frac{2\kappa}{n^2\Delta^2} \leq \frac{1}{2b^2}.
-\end{equation}
-
-For the second term, noting that $\Delta = \beta^T V_{\kappa} V_{\kappa}^T \beta/n$, we have
-\begin{align}
-\frac{\beta^T V_{\kappa} V_{\kappa}^T \beta/n^2}{\Delta^2} & = \frac{1}{n\Delta} \nonumber \\
-& \leq \frac{1}{2nt(b)} \nonumber \\
-& = \frac{1}{2b\sqrt{2\kappa}}, \label{eqn:spectral_type_II_error_pf2}
-\end{align}
-and combining~\eqref{eqn:spectral_type_II_error_pf1} and~\eqref{eqn:spectral_type_II_error_pf2} yields~\eqref{eqn:graph_spectral_type_II_error}.
-
-\subsection{Proof of Theorem~\ref{thm:sobolev_testing_rate_order1}}
-
-We want to apply Lemma~\ref{lem:fixed_graph_testing} to the case where $G = G_{n,r}$.
-In order to do this, we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$.
-In particular, we will show that the following statements each hold with probability at least $1 - c/b$ for sufficiently large $n$ (Here and in what follows, $c$ denotes a constant which is fixed in $n$ and $f$ and does not depend on $b$, but may depend on other fixed quantities such as $d$, $s$, etc.):
-\begin{enumerate}[label=(E\arabic*)]
-	\item 
-	\label{event:discrete_sobolev_norm_order1}
-	\textbf{Graph Sobolev norm:} For any $n^{-1/(2 + d)}\leq r \leq 1$,
-	\begin{equation}
-	\label{eqn:continuous_to_discrete_sobolev_norm_order1}
-	f^T L f \leq c \cdot b \cdot \norm{f}_{W^{1,2}(\Xset)}^2 \cdot n^{2} r^{d + 2} 
-	\end{equation}
-	\item 
-	\label{event:eigenvalue_tail_decay}
-	\textbf{Eigenvalue tail bound:} Let $t = n^{1/d}$, and let
-	\begin{equation*}
-	\overline{X} := \Bigl\{\frac{1}{t}(j_1,\ldots,j_d): k \in [t]^d\Bigr\}
-	\end{equation*} 
-	consist of $n$ evenly spaced grid points on $[0,1]^d$. The transportation distance between $X$ and $\ol{X}$ is
-	\begin{equation*}
-	d_{\infty}(X,\overline{X}) := \inf_{\pi} \sup\Bigl\{\norm{\pi(\ol{x}_i) - \ol{x}_i}: \ol{x}_i \in \ol{X}\Bigr\}
-	\end{equation*}
-	where the infimum is taken over all bijections $\pi: \ol{X} \to X$. 
-	There exists a constant $c$ such that for any $\kappa = 1,\ldots,n$ if $\max\{ 3 d_{\infty}(X,\ol{X}),n^{-1/d}\} \leq r \leq \kappa^{-2/(d(2 +d))}$, 
-	\begin{equation}
-	\label{eqn:eigenvalue_tail_bound}
-	\lambda_{\kappa} \geq c \cdot n r^{d + 2} \kappa^{2/d}
-	\end{equation}
-	Theorem~\ref{thm:slepcev_transport_distance} (Theorem 1.1 of \textcolor{red}{(Garcia Trillos and Slepcev)}) establishes that there exists a constant $c$ such that 
-	\begin{equation*}
-	d_{\infty}(X,\ol{X}) \leq c\biggl(\frac{\log n}{n}\biggr)^{1/d}
-	\end{equation*}
-	with probability $1  - O(1/n)$.
-Therefore in particular, the inequality~\eqref{eqn:eigenvalue_tail_bound} is satisfied for sufficiently large $n$ when $\kappa = n^{2d/(4 + d)}$ and $c(\log n/n)^{1/d} \leq r \leq n^{-4/((2+d)(4s + d))}$.
-	\item 
-	\label{event:l2_norm}
-	\textbf{Empirical norm of $f$:} There exists a constant $c_1$ such that if $\norm{f}_{\Leb^2(\Xset)} \geq c_1 \cdot b \cdot n^{-2/(4 + d)}$, and $d = 1,2$ or $3$,
-	\begin{equation}
-	\label{eqn:l2_to_empirical_norm}
-	\norm{f}_n^2 \geq \frac{1}{b} \cdot \norm{f}_{\Leb^2}^2
-	\end{equation}
-\end{enumerate}
-
-\subsubsection{Proof of \eqref{eqn:continuous_to_discrete_sobolev_norm}}
-
-The proof of~\eqref{eqn:continuous_to_discrete_sobolev_norm} follows from Lemma~\ref{lem:expected_first_order_seminorm}--which upper bounds the expected first order graph Sobolev seminorm-- and then an application of Markov's inequality.
-
-\subsubsection{Proof of \eqref{eqn:eigenvalue_tail_bound}}
-\label{subsubsec:eigenvalue_tail_bound_pf}
-
-We prove~\eqref{eqn:eigenvalue_tail_bound} by comparing $G_{n,r}$ to the tensor product of a $d$-dimensional lattice and a complete graph.
-The latter is a highly structured graph with known eigenvalues, which as we will see are sufficiently lower bounded for our purposes.
-
-Let $\wt{r} = r/(3(\sqrt{d} + 1)), M = (1/\wt{r})^d$ and $N = n\wt{r}^d$; assume without loss of generality that $M$ and $N$ are integers. Additionally for $m = M^{1/d}$ define
-\begin{equation*}
-\overline{Z} = \set{\frac{1}{m}(j_1,\ldots,j_d): j \in [m]^d}
-\end{equation*}
-to be the $M$ evenly spaced grid points over $[0,1]^d$.
-For a given $\overline{z}_j \in \overline{Z}$, we write $Q(z_j) = m^{-1}[j_1 - 1,j_1] \times \cdots \times m^{-1}[j_d - 1,j_d]$ for the cube of side length $1/m$ with $z_j$ at one corner.
-
-Consider the graph $H = (\overline{X}, E_H)$, where $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$ if
-\begin{equation*}
-\textrm{there exists}~\ol{z}_i, \ol{z}_j \in \ol{Z}~\textrm{such that}~\ol{x}_k \in Q(\ol{z}_i),~ \ol{x}_\ell \in Q(\ol{z}_j),~\textrm{and}~\norm{i - j}_1 \leq 1.
-\end{equation*}
-On the one hand $H \cong \ol{G}^M_d \otimes K_N$ where $\ol{G}^M_d$ is the $d$-dimensional lattice on $M$ nodes, and $K_N$ is the complete graph on $N$ nodes.
-On the other hand, we now show that when $\max\{3d_{\infty}(X,\ol{X}), n^{-1/d}\} \leq r$ then $G_{n,r} \succeq H$ as a result of the triangle inequality.
-If $(\ol{x}_k, \ol{x}_{\ell}) \in E(H)$, by definition there exist $\ol{z}_i, \ol{z}_j$ connected in $\ol{G}_d^M$ such that $\ol{x}_k \in Q(\ol{z}_i)$ and $\ol{x}_{\ell} \in Q(\ol{z}_j)$.
-This implies that $\ol{x}_k$ and $\ol{x}_{\ell}$ must themselves be close together, since
-\begin{equation*}
-\norm{\ol{x}_k - \ol{x}_{\ell}}_2 \leq \norm{\ol{x}_k - \ol{z}_{i}}_2 + \norm{\ol{z}_i - \ol{z}_j}_2 +  \norm{\ol{z}_{j} - \ol{x}_{\ell} }_2 \leq \wt{r}(1 + 2\sqrt{d}) = r/3.
-\end{equation*}
-Since we also assume $r/3 \geq d_{\infty}(X,\ol{X})$, another application of the triangle inequality gives
-\begin{equation*}
-\norm{\pi(\ol{x}_k) - \pi(\ol{x}_\ell)}_2 \leq \norm{\pi(\ol{x}_k) - \ol{x}_\ell}_2 + \norm{\ol{x}_k - \ol{x}_\ell}_2 \leq \norm{\ol{x}_{\ell} - \pi(\ol{x}_\ell)}_2 \leq r,
-\end{equation*}
-implying that $(\pi(\ol{x}_k), \pi(\ol{x}_{\ell})) \in E$ and consequently that $G_{n,r} \succeq \ol{G}^M_d \otimes K_N$.
-
-The eigenvalues of lattices and complete graphs are known to satisfy, respectively
-\begin{equation*}
-\lambda_k(\ol{G}^{M}_d) \geq \frac{k^{2/d}}{M^{2/d}}~\textrm{for $k = 0,\ldots,M - 1$},~~ \textrm{and}~\lambda_{j}(K_N) \geq N\1\{j > 0\}~\textrm{for $j = 0,\ldots,N-1$.}
-\end{equation*}
-and by standard facts regarding the eigenvalues of tensor product graphs, we have that the spectrum $\Lambda(H)$ satisfies
-\begin{equation*}
-\Lambda(H) = \set{N\lambda_k(\ol{G}^{M}_d) + M\lambda_j(K_N): \textrm{for $k = 0,\ldots,M - 1$ and $j = 0,\ldots,N-1$}}
-\end{equation*}
-For all $j = 1,\ldots,N-1$, we have that $M\lambda_j(K_N) = MN = n$. Therefore,
-\begin{align*}
-\lambda_{\kappa}(H) & \geq \{n \wedge N\lambda_{\kappa}(\ol{G}^{M}_d)\} \\
-& \geq \{n \wedge n\wt{r}^d\frac{\kappa^{2/d}}{M^{2/d}}\} \\
-& \geq \{n \wedge (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d}\} \\
-& \geq (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d},
-\end{align*}
-where the last inequality is satisfied since $r \leq \kappa^{-2/(d(d + 2))}$, completing the proof of~\eqref{eqn:eigenvalue_tail_bound}.
-
-\subsubsection{Proof of \eqref{eqn:l2_to_empirical_norm_order1}}
-The inequality~\eqref{eqn:l2_to_empirical_norm_order1} is implied by the more general Lemma~\ref{lem:empirical_norm_sobolev}, which we state in Section~\ref{subsec:sobolev_testing_rate_pf} and use to prove a similar inequality when $f \in W^{s,2}(\Xset)$ for general $s$.
-
-\subsubsection{Putting the pieces together}
-
-The final proof of Theorem~\ref{thm:sobolev_testing_rate_order1} proceeds by taking $s = 1$ and proceeding as in Section~\ref{subsubsec:sobolev_testing_rate_pf_conclusion}.
-
 \subsection{Proof of Theorem~\ref{thm:sobolev_testing_rate}}
-\label{subsec:sobolev_testing_rate_pf}
-
-We want to apply Lemma~\ref{lem:fixed_graph_testing} to the case where $G = G_{n,r}$.
+
+We want to apply Lemma~\ref{lem:fixed_graph_testing}, which analyzes testing over an arbitrary graph $G$, to the case where $G = G_{n,r}$.
 In order to do this, we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$.
@@
 	\item 
-	\label{event:l2_norm_order1}
-	\textbf{Empirical norm of $f$:} There exists a constant $c_1$ such that if $\norm{f}_{\Leb^2(\Xset)} \geq c_1 \cdot b \cdot n^{-2s/(4s + d)}$ and $4s > d$, then
-	\begin{equation}
-	\label{eqn:l2_to_empirical_norm_order1}
+	\label{event:l2_norm}
+	\textbf{Empirical norm of $f$:} There exists a constant $c_1$ such that if $\norm{f}_{\Leb^2(\Xset)} \geq c_1 \cdot b \cdot n^{-2s/(4s + d)}$ and $4s > d$,
+	\begin{equation}
+	\label{eqn:l2_to_empirical_norm}
 	\norm{f}_n^2 \geq \frac{1}{b} \cdot \norm{f}_{\Leb^2}^2
@@
 
-\subsubsection{Proof of~\eqref{eqn:eigenvalue_tail_bound}}
-See Section~\ref{subsubsec:eigenvalue_tail_bound_pf}.
+\subsubsection{Proof of~\eqref{eqn:eigenvalue_tail_bound_2}}
+
+We prove~\eqref{eqn:eigenvalue_tail_bound_2} by comparing $G_{n,r}$ to the tensor product of a $d$-dimensional lattice and a complete graph.
+The latter is a highly structured graph with known eigenvalues, which as we will see are sufficiently lower bounded for our purposes.
+
+Let $\wt{r} = r/(3(\sqrt{d} + 1)), M = (1/\wt{r})^d$ and $N = n\wt{r}^d$; assume without loss of generality that $M$ and $N$ are integers. Additionally for $m = M^{1/d}$ define
+\begin{equation*}
+\overline{Z} = \set{\frac{1}{m}(j_1,\ldots,j_d): j \in [m]^d}
+\end{equation*}
+to be the $M$ evenly spaced grid points over $[0,1]^d$.
+For a given $\overline{z}_j \in \overline{Z}$, we write $Q(z_j) = m^{-1}[j_1 - 1,j_1] \times \cdots \times m^{-1}[j_d - 1,j_d]$ for the cube of side length $1/m$ with $z_j$ at one corner.
+
+Consider the graph $H = (\overline{X}, E_H)$, where $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$ if
+\begin{equation*}
+\textrm{there exists}~\ol{z}_i, \ol{z}_j \in \ol{Z}~\textrm{such that}~\ol{x}_k \in Q(\ol{z}_i),~ \ol{x}_\ell \in Q(\ol{z}_j),~\textrm{and}~\norm{i - j}_1 \leq 1.
+\end{equation*}
+On the one hand $H \cong \ol{G}^M_d \otimes K_N$ where $\ol{G}^M_d$ is the $d$-dimensional lattice on $M$ nodes, and $K_N$ is the complete graph on $N$ nodes.
+On the other hand, we now show that when $\max\{3d_{\infty}(X,\ol{X}), n^{-1/d}\} \leq r$ then $G_{n,r} \succeq H$ as a result of the triangle inequality.
+If $(\ol{x}_k, \ol{x}_{\ell}) \in E(H)$, by definition there exist $\ol{z}_i, \ol{z}_j$ connected in $\ol{G}_d^M$ such that $\ol{x}_k \in Q(\ol{z}_i)$ and $\ol{x}_{\ell} \in Q(\ol{z}_j)$.
+This implies that $\ol{x}_k$ and $\ol{x}_{\ell}$ must themselves be close together, since
+\begin{equation*}
+\norm{\ol{x}_k - \ol{x}_{\ell}}_2 \leq \norm{\ol{x}_k - \ol{z}_{i}}_2 + \norm{\ol{z}_i - \ol{z}_j}_2 +  \norm{\ol{z}_{j} - \ol{x}_{\ell} }_2 \leq \wt{r}(1 + 2\sqrt{d}) = r/3.
+\end{equation*}
+Since we also assume $r/3 \geq d_{\infty}(X,\ol{X})$, another application of the triangle inequality gives
+\begin{equation*}
+\norm{\pi(\ol{x}_k) - \pi(\ol{x}_\ell)}_2 \leq \norm{\pi(\ol{x}_k) - \ol{x}_\ell}_2 + \norm{\ol{x}_k - \ol{x}_\ell}_2 \leq \norm{\ol{x}_{\ell} - \pi(\ol{x}_\ell)}_2 \leq r,
+\end{equation*}
+implying that $(\pi(\ol{x}_k), \pi(\ol{x}_{\ell})) \in E$ and consequently that $G_{n,r} \succeq \ol{G}^M_d \otimes K_N$.
+
+The eigenvalues of lattices and complete graphs are known to satisfy, respectively
+\begin{equation*}
+\lambda_k(\ol{G}^{M}_d) \geq \frac{k^{2/d}}{M^{2/d}}~\textrm{for $k = 0,\ldots,M - 1$},~~ \textrm{and}~\lambda_{j}(K_N) \geq N\1\{j > 0\}~\textrm{for $j = 0,\ldots,N-1$.}
+\end{equation*}
+and by standard facts regarding the eigenvalues of tensor product graphs, we have that the spectrum $\Lambda(H)$ satisfies
+\begin{equation*}
+\Lambda(H) = \set{N\lambda_k(\ol{G}^{M}_d) + M\lambda_j(K_N): \textrm{for $k = 0,\ldots,M - 1$ and $j = 0,\ldots,N-1$}}
+\end{equation*}
+For all $j = 1,\ldots,N-1$, we have that $M\lambda_j(K_N) = MN = n$. Therefore,
+\begin{align*}
+\lambda_{\kappa}(H) & \geq \{n \wedge N\lambda_{\kappa}(\ol{G}^{M}_d)\} \\
+& \geq \{n \wedge n\wt{r}^d\frac{\kappa^{2/d}}{M^{2/d}}\} \\
+& \geq \{n \wedge (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d}\} \\
+& \geq (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d},
+\end{align*}
+where the last inequality is satisfied since $r \leq \kappa^{-2/(d(d + 2))}$, completing the proof of~\eqref{eqn:eigenvalue_tail_bound_2}.
 
@@
 
-\subsubsection{Putting the pieces together}
-\label{subsubsec:sobolev_testing_rate_pf_conclusion}
-
-We note that for all possible values of $X \in \Xset^n$, under the null hypothesis $f = f_0 = 0$ and therefore $\beta = (f(x_1),\ldots,f(x_n)) = 0$ as well.
-Therefore by~\eqref{eqn:graph_spectral_type_I_error}, we have the following bound on Type I error:
-\begin{equation}
-\Ebb_{f_0}(\phi_{\mathrm{spec}}) = \mathbb{E}(\mathbb{E}_{\beta = 0}(\phi_{\spec}) | X) \leq \frac{1}{b^2}.
-\end{equation}
-
-Now, we bound Type II error under the assumption $f \in W_{\sigma}^{s,2}(\mathcal{X};R), p \in C^{s - 1}(\Xset;p_{\max})$ uniformly bounded away from $0$ and $\infty$ over $\Xset$, and 
-\begin{equation}
-\label{eqn:critical_radius_1}
-\norm{f}_{\Leb^2(\Xset)}^2 \geq \epsilon^2 = c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4s + d)}\} \cdot n^{-4s/(4s + d)}.
-\end{equation}
-Choosing $n^{-1/(2(s - 1) + d)}\leq r(n) \leq n^{-4/((2+d)(4s + d))}$, we may therefore apply our conclusions in Step 2; namely, that for every possible choice of $f \in W^{s,2}(\Xset;R)$ there exists a good set $\mathcal{E}_f \subseteq \Xset^n$ with $\Pbb(\mathcal{E}_f) \geq 1 - c/b$ such that each of \eqref{eqn:continuous_to_discrete_sobolev_norm}, \eqref{eqn:eigenvalue_tail_bound}, and \eqref{eqn:l2_to_empirical_norm_sobolev} hold for all $X \subseteq \mathcal{E}_f$.
-The choice $\kappa = (nR^2)^{2d/(4s + d)}$ balances the squared bias and variance terms on the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius}, and we have that for all $X \subseteq \mathcal{E}_f$
-\begin{align*}
-2b\sqrt{\frac{2\kappa}{n^2}} + \frac{f^T L^s f}{n\lambda_{\kappa}^s} & \leq 2bR^{2d/(4s + d)}n^{-4s/(4s+d)} + cbR^2\frac{n^{s} r^{s(d + 2)}}{\lambda_{\kappa}^s} \tag{by \eqref{eqn:continuous_to_discrete_sobolev_norm}} \\
-& \leq cbR^{2d/(4s + d)}n^{-4s/(4s+d)} \tag{by \eqref{eqn:eigenvalue_tail_bound}} \\
-& \leq \frac{1}{b}\norm{f}_{\Leb^2(\Xset)}^2 \\
-& \leq \frac{1}{n}\sum_{i = 1} \bigl[f(x_i)\bigr]^2. \tag{by \eqref{eqn:l2_to_empirical_norm_sobolev}}
-\end{align*}
-where the last two inequalities follow for a suitably large choice of $c_1$ in~\eqref{eqn:critical_radius_1}.
-We conclude that for all $X \subseteq \mathcal{E}_f$, the inequality \eqref{eqn:fixed_graph_testing_critical_radius} is satisfied with respect to $\beta = (f(x_1),\ldots,f(x_n))$ and $G = G_{n,r}$.
-As a result the worst-case Type II error is bounded
-\begin{equation*}
-\sup_{\substack{f \in W^{s,2}(\Xset;R), \\ \norm{f}_{\Leb^2(\Xset)} \geq \epsilon}}\mathbb{E}_{f}(1 - \phi_{\spec}) \leq \sup_{\substack{f \in W^{s,2}(\Xset;R), \\ \norm{f}_{\Leb^2(\Xset)} \geq \epsilon}} \mathbb{E}\bigl[\mathbb{E}_{\beta}(1 - \phi_{\spec}|X \in \mathcal{E}_f)\bigr] + \frac{c}{b} \leq \frac{3 + c}{b},
-\end{equation*}
-completing the proof of Theorem~\ref{thm:sobolev_testing_rate} upon proper choice of constants $c_1,c_2$ in that Theorem.
-
-\subsection{Proof of Proposition~\ref{prop:L4_testing_rate}}
-
-Let $T_{\mathrm{mean}} = \frac{1}{n}\sum_{i = 1}^{n} y_i^2$. The expectation of $T_{\mathrm{mean}}$ is
-\begin{equation*}
-\Ebb(T_{\mathrm{mean}}) = \mathbb{E}(f^2(x_1)) + 1,
-\end{equation*}
-and the variance can be upper bounded
-\begin{equation*}
-\Var(T_{\mathrm{mean}}) \leq \frac{1}{n}(3 + p_{\max} \norm{f}_{\Leb^4}^4 + p_{\max}\norm{f}_{\Leb^2}^2) = \frac{c}{n}.
-\end{equation*}
-Since $\mathbb{E}(f^2(x_1)) \geq p_{\min} \epsilon^2$, when $\epsilon^2 \gtrsim b\cdot n^{-1/2}$ we can apply Chebyshev's inequality to obtain the claimed result.
-
-\subsection{Proof of Theorem~\ref{thm:twosample_sobolev_testing_rate}}
-
-As mentioned previously, to prove Theorem~\ref{thm:twosample_sobolev_testing_rate}, we relate the two sample model to the following regression model: we observe
-$X = \{x_1,\ldots,x_n\} \sim \mu$ (where we recall $\mu = (P + Q)/2$), and associated labels
-\begin{equation*}
-a_i = 
-\begin{cases}
-1, & \textrm{with probability}~ \frac{p(x_i)}{p(x_i) + q(x_i)} \\
--1, & \textrm{with probability}~ \frac{q(x_i)}{p(x_i) + q(x_i)}
-\end{cases}
-\end{equation*}
-where $a_i$ is conditionally independent of $x_j,a_j$ given $x_i$.
-
-\textcolor{red}{TODO:} Complete the above argument.
-
-Our analysis will trace a similar path to the proof of Theorem~\ref{thm:sobolev_testing_rate}, proceeding according to the following steps:
-\begin{enumerate}
-	\item We upper bound the testing error of our eigenvector projection test statistic when $a$ are viewed as random responses defined over the vertices of a fixed graph $G$.
-This upper bound will hold whenever certain functionals on the graph $G$ are themselves bounded.
-One of these functionals will be a measure of eigenvector incoherence.
-	\item We analyze the behavior of these functionals with respect to the random graph $G_{n,r}$, and bound them with high probability.
-	\item We condition on $X \subseteq \mathcal{E}$, where $\mathcal{E}$ is a high-probability set over which the relevant functionals on $G_{n,r}$ are bounded.
-We conclude that the upper bound on testing error derived in our first step holds whenever $X \subseteq \mathcal{E}$.
-\end{enumerate}
-
-\subsubsection{Step 1: Testing error on a fixed graph}
-
-Let $G = (V,E)$ be a graph over vertices $V = \set{v_1,\ldots,v_n}$, and let $\beta^{(p)} = (\beta_1^{(p)},\ldots,\beta_n^{(p)}) \in \Reals^n$, $\beta^{(q)} = (\beta_1^{(q)},\ldots,\beta_n^{(q)}) \in \Reals^n$ be non-negative signals over the vertices $V$.
-We observe labels $a = (a_1,\ldots,a_n)$ according to the model
-\begin{equation*}
-a_i = 
-\begin{cases*}
-1, & \textrm{with probability $\frac{\betap_i}{\betap_i + \betaq_i}$} \\
--1, & \textrm{with probability $\frac{\betaq_i}{\betap_i + \betaq_i}$}
-\end{cases*}
-\end{equation*}
-Letting $L = VSV^T$ be the spectral decomposition of the Laplacian $L$ of $G$, our graph spectral test statistic is 
-\begin{equation*}
-T_{\spec}^{(2)} = \frac{1}{n}\sum_{k = 1}^{\kappa} \left(\sum_{i = 1}^{n} v_i a_i\right)^2
-\end{equation*}
-where $\kappa$ is a tuning parameter. The resulting test we will use is
-\begin{equation*}
-\phi_{\spec}^{(2)} = \1\{T_{\spec}^{(2)} \geq \frac{\kappa}{n} + t(b)\},~~\textrm{where}~ t(b) = b\sqrt{\frac{2\kappa}{n^2}}~~\textrm{for $b \geq 1$.}
-\end{equation*}
-
-Let $\Pi_{\max}(\kappa;G)$ be a measure of the incoherence of the eigenvectors $v_1,\ldots,v_\kappa$, given by
-\begin{equation*}
-\Pi_{\max}(\kappa;G) := \max_{i = 1,\ldots,n} \left\{\sum_{k = 1}^{\kappa} v_{k,i}^2\right\}
-\end{equation*}
-In Lemma~\ref{lem:twosample_fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}^{(2)}$.
-Our bound on the Type II error will be stated as a function of $\Pi_{\max}(\kappa;G)$ as well as the $\kappa$th eigenvalue $s_{\kappa}$ and the smoothness functional $S_2(\beta;G)$.
-We use the notation $\varDelta^{(p,q)} = (\betap - \betaq)/(\betap + \betaq)$.
-
-\begin{lemma}
-	\label{lem:twosample_fixed_graph_testing}
-	Let $1 \leq \kappa \leq n$ be an integer.
-	\begin{enumerate}
-		\item \textbf{Type I error:} Under the null hypothesis $\betap = \betaq$, the Type I error of $\phi_{\spec}^{(2)}$ is upper bounded
-		\begin{equation}
-		\label{eqn:twosample_graph_spectral_type_I_error}
-		\mathbb{E}_{\betap,\betap}(\phi_{\spec}^{(2)}) \leq \frac{1}{b^2}.
-		\end{equation}
-		\item \textbf{Type II error:} Suppose $\Pi_{\max}(\kappa;G) \leq 1$. Then for any $b$, $\betap$ and $\betaq$ such that
-		\begin{equation}
-		\label{eqn:twosample_fixed_graph_testing_critical_radius}
-		\frac{1}{n} \sum_{i = 1}^{n} (\varDelta_i^{(p,q)})^2 \geq \frac{1}{1 - \Pi_{\max}(\kappa;G)}\left(2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_2(\varDelta;G)}{ns_{\kappa}}\right)
-		\end{equation}
-		the Type II error of $\phi_{\spec}^{(2)}$ is upper bounded,
-		\begin{equation}
-		\label{eqn:twosample_graph_spectral_type_II_error}
-		\mathbb{E}_{\betap,\betaq}(1 - \phi_{\spec}^{(2)}) \leq \frac{3}{b}.
-		\end{equation}
-	\end{enumerate}
-\end{lemma}
-
-\paragraph{Proof of Lemma~\ref{lem:twosample_fixed_graph_testing}.}
-
-To prove Lemma~\ref{lem:twosample_fixed_graph_testing} we will compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}^{(2)}$.
-The inequalities \eqref{eqn:twosample_graph_spectral_type_I_error} and \eqref{eqn:twosample_graph_spectral_type_II_error} can then be computed using Chebyshev's inequality in a manner very similar to the proof of Lemma~\ref{lem:fixed_graph_testing}, and we omit the details.
-Let $w_i := a_i - \vardeltapq_{i}$, and note that $\Ebb(w_i) = 0$ and
-\begin{equation*}
-\Ebb(w_i^2) = \frac{\betap_i \betaq_i}{(\betap_i + \betaq_i)^2} = 1 - (\vardeltapq_{i})^2,~~ \Ebb(w_i^3) = \frac{1}{2}(1 - (\vardeltapq_{i})^2)\vardeltapq_{i}
-\end{equation*}
-
-\vspace{.2 in}
-
-\textit{Mean of $T_{\spec}^{(2)}$}:
-
-We have
-\begin{align*}
-\Ebb(T_{\spec}^{(2)}) & = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{\vardeltapq}{v_k}^2 + \Ebb\bigl(\dotp{w}{v_k}^2 + 2 \dotp{w}{v_k} \dotp{\vardeltapq}{v_k}\bigr)\right) \\
-& = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{1 - (\vardeltapq)^2}{v_k^2} + \dotp{\vardeltapq}{v_k}\right) \\
-& \geq \frac{\kappa}{n} + \left(\sum_{k = 1}^{\kappa}\dotp{\vardeltapq}{v_k} - \dotp{(\vardeltapq)^2}{v_k^2}\right)
-\end{align*}
-When $\vardeltapq = 0$, this equals $\kappa/n$. Otherwise, we have the following upper bound:
-\begin{align*}
-\sum_{k = 1}^{\kappa} \dotp{(\vardeltapq)^2}{v_k^2}  & = \sum_{i = 1}^{n} (\vardeltapq_{i})^2 \left(\sum_{k = 1}^{\kappa} v_{k,i}^2 \right)\\
-& \leq \Pi_{\max}(\kappa,G) \cdot \sum_{i = 1}^{n} (\vardeltapq_{i})^2 = \Pi_{\max}(\kappa,G) \cdot \norm{\vardeltapq}_2^2.
-\end{align*}
-Additionally, $\sum_{k = 1}^{\kappa}\dotp{\vardeltapq}{v_k} \geq \norm{\vardeltapq}_2^2 - S_2(\beta;G)/s_{\kappa}$ by reasoning given in the proof of Lemma~\ref{lem:fixed_graph_testing}, and as a result
-\begin{equation*}
-\Ebb(T_{\spec}^{(2)}) \geq \frac{\kappa}{n} -  \frac{\bigl(1 - \Pi_{\max}(\kappa;G)\bigr)}{n} \norm{\vardeltapq}_2^2 - \frac{S_2(\beta;G)}{ns_{\kappa}}
-\end{equation*}
-
-\vspace{.2 in}
-\textit{Variance of $T_{\spec}^{(2)}$:}
-
-We have
-\begin{align}
-\Var(T_{\spec}^{(2)}) & = \frac{1}{n^2}\Var(2\dotp{V_{\kappa}V_{\kappa}^T\vardeltapq}{w} + \dotp{V_{\kappa}V_{\kappa}^Tw}{w}) \nonumber \\
-& = \frac{1}{n^2}\Bigl\{4\underbrace{\Var(\dotp{V_{\kappa}V_{\kappa}^T\vardeltapq}{w})}_{=: V_1} + 2\underbrace{\Cov(\dotp{V_{\kappa}V_{\kappa}^T\vardeltapq}{w},\dotp{V_{\kappa}V_{\kappa}^Tw}{w})}_{=: K_1} + \underbrace{\Var(\dotp{V_{\kappa}V_{\kappa}^Tw}{w})}_{=:V_2}\Bigr\}, \label{eqn:var}
-\end{align}
-and we now upper bound each of the three terms on the right hand side of the previous display.
-
-\textbf{Upper bound on $V_1$:}
-Let $\Sigma := \Cov(w)$ be the covariance matrix of $w$. Noting that $\Sigma \preceq I$, we have
-\begin{equation}
-\label{eqn:var1}
-\Var(\dotp{V_{\kappa}V_{\kappa}^T\vardeltapq}{w}) =(\vardeltapq)^T V_{\kappa}V_{\kappa}^T \Sigma V_{\kappa}V_{\kappa}^T \vardeltapq \leq \norm{V_{\kappa}V_{\kappa}^T\vardeltapq}^2.
-\end{equation}
-
-\textbf{Upper bound on $K_1$:}
-Noting that $\Ebb(\dotp{V_{\kappa}V_{\kappa}^T \vardeltapq}{w}) = 0$, we have that
-\begin{align}
-K_1 & = \Ebb\left[\dotp{V_{\kappa}V_{\kappa}^T \vardeltapq}{w} \dotp{V_{\kappa}V_{\kappa}^T w}{w}\right] \nonumber \\
-& = \sum_{i = 1}^{n} \sum_{j = 1}^{n} \sum_{i' = 1}^{n} \Ebb\left[w_i w_j (V_{\kappa}V_{\kappa}^T)_{ij} w_{i'} (V_{\kappa}V_{\kappa}^T\vardeltapq)_i\right] \nonumber \\
-& = \sum_{i = 1}^{n} \Ebb\left[w_i^3\right] (V_{\kappa}V_{\kappa}^T)_{ii} (V_{\kappa}V_{\kappa}^T \vardeltapq)_i \nonumber \\
-& = \frac{1}{2}\sum_{i = 1}^{n} (1 - (\vardeltapq_{i})^2)(\vardeltapq_{i}) (V_{\kappa} V_{\kappa}^T)_{ii}  (V_{\kappa}V_{\kappa}^T\vardeltapq)_i \nonumber \\
-& \leq \norm{V_{\kappa}V_{\kappa}^T \vardeltapq} \cdot \left(\sum_{i = 1}^{n} (\vardeltapq_{i})^2 (V_{\kappa} V_{\kappa}^T)_{ii}^2\right)^{1/2} \nonumber \\
-& \leq \Pi(\kappa,G) \cdot \norm{V_{\kappa}V_{\kappa}^T \vardeltapq} \cdot \norm{\vardeltapq}  \label{eqn:var2} 
-\end{align}
-
-\textbf{Upper bound on $V_2$:}
-$V_2$ is a variance of a sum, which we re-express as the sum of covariances:
-\begin{align*}
-V_2 & = \Var(\dotp{V_{\kappa}V_{\kappa}^T w}{w}) \\
-& = \sum_{i,j,i',j' = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ij} (V_{\kappa}V_{\kappa}^T)_{i'j'} \Cov(w_i w_j, w_{i'} w_{j'}).
-\end{align*}
-This covariance will be non-zero only when $i = i' \neq j = j'$, $i = j' \neq j = i'$, or $i = i' = j = j'$, and therefore
-\begin{align}
-V_2 & = 2\sum_{i = 1}^{n}\sum_{j = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ij}^2 \Var(w_iw_j) + \sum_{i = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ii}^2 \Var(w_i^2) \nonumber\\
-& \leq 2\sum_{i = 1}^{n}\sum_{j = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ij}^2 + \sum_{i = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ii}^2 \leq 3 \mathrm{tr}\bigl((V_{\kappa}V_{\kappa}^T)^2\bigr) = 3\kappa.
-\label{eqn:var3}
-\end{align}
-
-Now, plugging \eqref{eqn:var1}, \eqref{eqn:var2}, and \eqref{eqn:var3} back into \eqref{eqn:var}, we obtain
-\begin{equation}
-\label{eqn:var_5}
-\Var(T_{\spec}^{(2)}) \leq \frac{1}{n^2}\left\{4\norm{V_{\kappa}V_{\kappa}^T \vardeltapq}^2 + \Pi(\kappa;G) \cdot \norm{V_{\kappa}V_{\kappa}^T \vardeltapq} \cdot \norm{\vardeltapq} + 3\kappa \right\}.
-\end{equation}
-
-\subsubsection{Step 2: Bounding neighborhood graph functionals}
-
-Note that since $p,q \in W^{1,2}(\mathcal{X};R)$, we have $f := p - q \in W^{1,2}(\mathcal{X};2R)$.
-Therefore the bounds~\eqref{eqn:continuous_to_discrete_sobolev_norm} and \eqref{eqn:l2_to_empirical_norm} apply to $S_2(\varDelta,G_{n,r})$ and $(1/n)\sum_{i = 1}^{n}\varDelta_i^2$, respectively.
-The bound~\eqref{eqn:eigenvalue_tail_bound} continues to apply to $s_{\kappa}$.
-What remains is to upper bound the incoherence parameter
-\begin{equation}
-\label{eqn:pi_max}
-\Pi_{\max}(\kappa,G_{n,r}) \leq \frac{1}{2}~~\textrm{when}~ \kappa = n^{2/5}~\textrm{and}~ r = \log^a n \cdot \left(\frac{\log n}{n}\right)
-\end{equation}
-with probability $1 - o(1)$ as $n \to \infty$.
-
-Our proof of~\eqref{eqn:pi_max} will proceed by relating $G_{n,r}$ to a graph which satsifies a strict notion of incoherence, the chain graph.
-When $G = \overline{G}$, the incoherence parameter $\Pi_{\max}$ is upper bounded
-\begin{equation}
-\Pi_{\max}(\kappa,\overline{G}) \leq \frac{\kappa}{n}.
-\end{equation}
-To prove that the weaker bound \eqref{eqn:pi_max} holds with respect to $G_{n,r}$, we will show that the two graphs $\overline{G}$ and $G_{n,r}$ are $\delta$-spectrally similar, meaning
-\begin{equation}
-\label{eqn:spectral_similarity}
-(1 - \delta_n) x^T L x \leq x^T \overline{L} x \leq (1 + \delta_n) x^T L x~~\textrm{for all}~x \in \Reals^n, \delta_n = c \cdot (\log n)^{h(a,1) + da}\cdot \frac{\log n}{n}
-\end{equation}
-with probability $1 - o(1)$ as $n \to \infty$.
-The proof of~\eqref{eqn:spectral_similarity} relies heavily on bounding the transport distance between $X$ and $\overline{X}$, i.e.
-\eqref{eqn:slepcev_transport_distance}.
-Then the following technical Lemma along with \eqref{eqn:spectral_similarity} will imply \eqref{eqn:pi_max}.
-\begin{lemma}
-	\label{lem:pi_max_pf_1}
-	Let $G$ satisfy~\eqref{eqn:spectral_similarity} for a given $0 \leq \delta < 1$. Then for any $R \gtrsim n \delta^{1/2} (s_{n}^{1/2} \vee 1)$ and $k \leq n/8$,
-	\begin{equation*}
-	v_{k,i}^2 \lesssim \frac{1}{n}\left(R + \frac{n^4 \delta^2 s_{n}^2}{R^3}\right) \quad \textrm{for every $i = 1,\ldots,n$.}
-	\end{equation*} 
-\end{lemma}
-
-\textcolor{red}{TODO}: Explain how Lemma~\ref{lem:pi_max_pf_1} implies~\eqref{eqn:pi_max}.
-
-At a high level, Lemma~\ref{lem:pi_max_pf_1} is proved through repeated applications of the Davis-Kahan Theorem.
-The full proof is long and we delay presenting it until Section~\ref{sec:technical_lemma_proofs}.
-
-\subsubsection{Step 3: Conclusion}
-
-With Lemma~\ref{lem:twosample_fixed_graph_testing} as well as \eqref{eqn:continuous_to_discrete_sobolev_norm}, \eqref{eqn:eigenvalue_tail_bound}, \eqref{eqn:l2_to_empirical_norm} and \eqref{eqn:pi_max} in hand, the proof of Theorem~\ref{thm:twosample_sobolev_testing_rate} follows by similar reasoning to the conclusion of Theorem~\ref{thm:sobolev_testing_rate}.
-
-\section{Proofs of Technical Results}
+\subsection{Proof of Lemma~\ref{lem:roughness_functional_expectation_sobolev}}
 \label{sec:technical_lemma_proofs}
 
-Here we give in full detail the proofs of Lemma~\ref{lem:roughness_functional_expectation_sobolev}, \eqref{eqn:spectral_similarity}, and Lemma~\ref{lem:pi_max_pf_1}.
-
-\subsection{Proof of Lemma~\ref{lem:roughness_functional_expectation_sobolev}}
 To simplify exposition, we introduce the iterated difference operator, defined recursively as
@@
 
-\subsection{Proof of~\eqref{eqn:spectral_similarity}}
-
-Assuming~\eqref{eqn:slepcev_transport_distance} holds and $r$ is chosen according to~\eqref{eqn:pi_max}, we have already shown that the upper bound in~\eqref{eqn:spectral_similarity} holds for $\delta = 0$ (see the proof of ~\eqref{eqn:eigenvalue_tail_bound}).
-Now we need only to show the lower bound, which we will do following these steps:
-\begin{enumerate}[(i)]
-	\item First, we consider the case of two fixed graphs $G$ and $\wt{G}$, and establish a Poincare inequality.
-In other words, we show that if edges in $G$ can be mapped to short, non-overlapping paths in $\wt{G}$, then the Laplacian matrix $L$ cannot be much larger $\wt{L}$, in the usual sense of partial ordering of matrices.
-	\item Then, we treat the case of $G = G_{n,r}$ and $\wt{G} = \overline{G}$.
-We assume that~\eqref{eqn:slepcev_transport_distance} holds, and exhibit such a mapping between edges in $G_{n,r}$ and paths in $\overline{G}$.
-The resulting Poincare inequality will imply~\eqref{eqn:spectral_similarity}.
+\section{Additional Results}
+
+\subsection{Fixed Graph Testing}
+
+\begin{lemma}
+	\label{lem:fixed_graph_testing}
+	Let $1 \leq \kappa \leq n$ be an integer.
+	\begin{enumerate}
+		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}$ is upper bounded
+		\begin{equation}
+		\label{eqn:graph_spectral_type_I_error}
+		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{1}{b^2}.
+		\end{equation}
+		\item \textbf{Type II error:} For any $b \geq 1$ and $\beta$ such that
+		\begin{equation}
+		\label{eqn:fixed_graph_testing_critical_radius}
+		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{\beta^T L^s \beta}{n\lambda_{\kappa}^s}
+		\end{equation}
+		the Type II error of $\phi_{\spec}$ is upper bounded,
+		\begin{equation}
+		\label{eqn:graph_spectral_type_II_error}
+		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{3}{b}.
+		\end{equation}
 	\end{enumerate}
-In step (ii) we assume that~\eqref{eqn:slepcev_transport_distance} holds.
-Since \eqref{eqn:slepcev_transport_distance} holds with probability at least $1 - c/n$, this will imply that~\eqref{eqn:spectral_similarity} holds with at least the same probability.
-
-\subsubsection{Step 1: Poincare inequality}
-
-Let $G = (V,E)$ and $\wt{G} = (V,\wt{E})$ be two graphs over a common vertex set $V$. The set $\mathcal{P}_{\wt{G}}$ consists of all paths $P$ over $\wt{G}$, meaning all tuples
-\begin{equation*}
-P = (\wt{e}_1,\wt{e}_2,\ldots,\wt{e}_m),~~m \in \mathbb{N}~, \wt{e}_j \in \wt{E}~ \forall~ j \in [m], (\wt{e}_j)_2 = (\wt{e}_{j + 1})_1 \forall~j \in [m - 1].
-\end{equation*}
-Let $\gamma:E \to \mathcal{P}_{\wt{G}}$ be a mapping from edges in $G$ to paths in $\wt{G}$. The maximum path length $M(\gamma)$ is defined
-\begin{equation*}
-M(\gamma) = \max_{e \in E} \abs{\gamma(e)}
-\end{equation*} 
-and the bottleneck $b(\gamma)$ is defined
-\begin{equation*}
-b(\gamma) = \max_{\wt{e} \in \wt{E}} \abs{\{e \in E:  \wt{e} \in \gamma(e)\}}.
-\end{equation*}
-As mentioned previously, the Laplacian matrix $L$ cannot be much larger than $\wt{L}$ if edges in $G$ can be mapped to short, non-overlapping paths in $\wt{G}$.
-Formally, we will show the following: for any $\gamma: E \to \mathcal{P}_{\wt{E}}$,
-\begin{equation}
-\label{eqn:poincare_inequality}
-x^T L x \leq M(\gamma)\cdot b(\gamma) \cdot (x^T \wt{L} x)~\forall{x \in \Reals^n}
-\end{equation}
-
-\paragraph{Proof of~\eqref{eqn:poincare_inequality}:}
-
-Letting $c \in \Reals$, we will use the notation $G \preceq c \cdot  \wt{G}$ as shorthand for
-\begin{equation*}
-x^T L x \leq c \cdot  x^T \wt{L} x,~~\textrm{for all $x \in \Reals^n$}.
-\end{equation*}
-
-Let $G_e = (V, \set{e})$ and $P_e = (V, \set{\widetilde{e}: \widetilde{e} \in \gamma(e)})$ be the graphs associated with $e$ and $\gamma(e)$, respectively. By Lemma \ref{lem: path_poincare}, we have
-\begin{equation*}
-G_{e} \preceq \abs{P_e} P_e
-\end{equation*}
-Summing over all $e \in E_G$, we obtain
-\begin{align*}
-G & \preceq \sum_{e \in E_G} \abs{P_e} P_e \\
-& \preceq \max_{e \in E_G} \abs{\gamma(e)} \sum_{e \in E_G} P_e \\
-& \preceq \max_{e \in E_G} \abs{\gamma(e)} b_{\gamma}\cdot \widetilde{G}
-\end{align*}
-
-\subsubsection{Step 2: Mapping $E_{G_{n,r}}$ to $\mathcal{P}_{\overline{G}}$}
-We assume there exists a mapping $\pi:\overline{X} \to X$ which satisfies \eqref{eqn:slepcev_transport_distance}.
-We will apply the Poincare inequality to $\pi^{-1}(G_{n,r})$, the isomorphism of $G_{n,r}$ induced by the mapping of vertices $\pi^{-1}: X \to \overline{X}$.
-We will map a given $(\overline{x}_{k},\overline{x}_{\ell}) \in \pi^{-1}(E)$ to a shortest path $P \in \mathcal{P}_{\overline{G}}$ (measured by Manhattan distance) between $\overline{x}_k$ and $\overline{x}_{\ell}$.
-Recalling the notation $t = n^{1/d}$ and letting
-\begin{equation*}
-s_i = \frac{1}{t}(\textrm{sign}(k_i - \ell_i),0,\ldots,0),~~\textrm{for $i = 1,\ldots,d$,}
-\end{equation*}
-our mapping $\gamma: \pi^{-1}(E) \to \mathcal{P}_{\overline{G}}$ is given by
-\begin{align*}
-\gamma((\overline{x}_{k},\overline{x}_{\ell})) = \bigl(& (\overline{x}_k, \overline{x}_{k} + s_1), (\overline{x}_{k} + s_1,\overline{x}_{k} + 2s_1),\ldots,(\overline{x}_{k} + (\abs{k_1 - \ell_1} - 1)s_1, \overline{x}_{k} + \frac{1}{t}(k_1 - \ell_1)) \\
-& (\overline{x}_{k} + \frac{1}{t}(k_1 - \ell_1), \overline{x}_{k} + \frac{1}{t}(k_1 - \ell_1) + s_2),\ldots,(\overline{x}_{k} + \frac{1}{t}(k_1 - \ell_1) + (\abs{k_2 - \ell_2} - 1)s_2, \overline{x}_{k}  + \frac{1}{t}((k_1 - \ell_1) + (k_2 - \ell_2)) \\
-& \vdots \\
-& (\overline{x}_{k} + \frac{1}{t}\sum_{i = 1}^{d - 1}(k_i - \ell_i) + s_d,\overline{x}_{k} + \frac{1}{t}\sum_{i = 1}^{d - 1}(k_i - \ell_i) + 2s_d), \ldots, (\overline{x}_k + \frac{1}{t}\sum_{i = 1}^{d - 1}(k_i - \ell_i) + (\abs{k_d - \ell_d} - 1)s_d, \overline{x}_{\ell})\bigr)
-\end{align*}
-As a result of~\eqref{eqn:slepcev_transport_distance} we obtain the following bounds on $M(\gamma)$ and $b(\gamma)$:
-\begin{align}
-M(\gamma) & \leq c n^{1/d} \left(\left(\frac{\log n}{n}\right)^{1/d} + r\right) \label{eqn:maximum_path_length}\\
-b(\gamma) & \leq c \left(n^{1/d} \left(\left(\frac{\log n}{n}\right)^{1/d} + r\right)\right)^{2d}, \label{eqn:bottleneck}
-\end{align}
-which we now prove.
-
-\paragraph{Proof of~\eqref{eqn:maximum_path_length}:}
-
-Suppose $(\overline{x}_k, \overline{x}_{\ell}) \in \pi^{-1}(E)$. Then
-\begin{align*}
-\gamma(\overline{x}_k, \overline{x}_{\ell})) = \norm{\overline{x}_k - \overline{x}_{\ell}}_1 & \leq \sqrt{d} \norm{\overline{x}_k - \overline{x}_{\ell}}_2 \\
-& \leq \sqrt{d}\bigl(\norm{\overline{x}_k - \pi(\ol{x}_{k})}_2 + \norm{\pi(\ol{x}_k) - \pi(\ol{x}_{\ell})}_2 + \norm{\overline{x}_\ell - \pi(\ol{x}_{\ell})}_2\bigr) \\
-& \leq \sqrt{d}\left(\norm{\pi(\ol{x}_k) - \pi(\ol{x}_{\ell})}_2 + 2c\left(\frac{\log n}{n}\right)^{1/d}\right) \tag{by \eqref{eqn:transport_distance}} \\
-& \leq \sqrt{d}\left(r + 2c\left(\frac{\log n}{n}\right)^{1/d}\right) \tag{since $(\pi(\ol{x}_k),\pi(\ol{x}_{\ell})) \in E$}
-\end{align*}
-
-\paragraph{Proof of~\eqref{eqn:bottleneck}:}
-
-Suppose $(\overline{x}_k, \overline{x}_{\ell}) \in \gamma(e)$ for some $e = (\overline{x}_i, \overline{x}_j) \in \pi^{-1}(E)$.  Then, by the construction of $\gamma$,
-\begin{equation}
-\label{eqn:bottleneck_pf_1}
-\norm{\overline{x}_k - \overline{x}_i}_2 \wedge \norm{\overline{x}_\ell - \overline{x}_j}_2  \leq \norm{\overline{x}_j - \overline{x}_i}_2
-\end{equation}
-As shown in the preceding paragraph, assuming~\eqref{eqn:transport_distance} we have
-\begin{equation*}
-\norm{\overline{x}_j - \overline{x}_i}_2 \leq r + 2c\left(\frac{\log n}{n}\right)^{1/d}
-\end{equation*}
-and therefore by~\eqref{eqn:bottleneck_pf_1},
-\begin{align*}
-\abs{\set{e \in E: (\overline{x}_k, \overline{x}_{\ell}) \in \gamma(e)}} & \leq \abs{\set{\overline{x}_i: \norm{\overline{x}_k - \overline{x}_i}_2 \leq r + 2c\left(\frac{\log n}{n}\right)^{1/d}}} \cdot \abs{\set{\overline{x}_i: \norm{\overline{x}_k - \overline{x}_i}_2 \leq r + 2c\left(\frac{\log n}{n}\right)^{1/d}}} \\
-& \leq \left(n^{1/d}\left( r + 2c\left(\frac{\log n}{n}\right)^{1/d}\right)\right)^2.
-\end{align*}
-
-Putting the pieces together, we have that when $d = 1$ and $r = (\log n)^a (\log n/n)$, then
-\begin{equation*}
-G_{n,r} \preceq c \cdot (\log n)^{(1/d + a) + 2 + 2da}\cdot \overline{G}
-\end{equation*}
-with probability at least $1 - o(1)$.
-
-\subsection{Proof of Lemma~\ref{lem:pi_max_pf_1}}
-
-Write $\overline{L} = \overline{V}\overline{S} \overline{V}^T$ for the spectral decomposition of $\overline{L}$. The proof of Lemma~\ref{lem:pi_max_pf_1} will proceed according to the following steps.
-\begin{enumerate}
-	\item We show that the incoherence $\max_{i} v_{k,i}^2$ can be upper bounded
-	\begin{equation}
-	\label{eqn:incoherence_proof_1}
-	\max_{i} v_{k,i}^2 \leq \frac{1}{n} \left(\sum_{j = 1}^{n} \abs{\dotp{v_k}{\overline{v}_j}}\right)^2
-	\end{equation}
-	replacing the norm $\norm{v_k^2}_{\infty}$ by the inner products $\dotp{v_k}{\overline{v}_j}$.
-	\item Let $I(k,r) = \set{j \geq 0: \abs{j - k} \leq r}$. Using Davis-Kahan, we establish that the following upper bounds
-	\begin{equation}
-	\label{eqn:incoherence_proof_2}
-	\sum_{j \not\in I(k,r)}^{n} (\dotp{v_k}{\overline{v}_j})^2 \leq 400 \frac{n^4 \delta^2 s_{\max}^2}{R^4}
-	\end{equation}
-	hold for each $k \leq n/8$ and any $r \geq 8k\sqrt{\delta}$. 
-	\item We carefully upper bound the $L_1$ norm present in \eqref{eqn:incoherence_proof_1} given the various bounds on $L_2$ norm we've established in \eqref{eqn:incoherence_proof_2}.
-\end{enumerate}
-
-\paragraph{Step 1: Proof of~\eqref{eqn:incoherence_proof_1}.}
-
-We can re-express $v_k$ in the basis $(\overline{v}_j)$ as
-\begin{equation*}
-v_k = \sum_{j = 1}^{n} \dotp{v_k}{\overline{v}_j} \overline{v}_j,
-\end{equation*}
-whereupon the bound~\eqref{eqn:incoherence_proof_1} follows from the incoherence property of the chain
-\begin{equation*}
-\max_{i,k = 1,\ldots,n} \overline{v}_{k,i}^2 \leq \frac{1}{n}.
-\end{equation*}
-\paragraph{Step 2: Proof of~\eqref{eqn:incoherence_proof_2}.}
-
-Let $\widetilde{L} = L + H$. By Davis Kahan, we have that
-\begin{equation}
-\label{eqn:davis_kahan}
-\sum_{j \not\in I(k,r)} (\dotp{v_k}{\overline{v}_j})^2 \leq \left(\frac{\norm{H}_{\textrm{op}}}{\min{\abs{\overline{s}_j - s_k}:j \in I(k,r)}}\right)^2
-\end{equation}
-To upper bound the numerator, we use~\eqref{eqn:spectral_similarity} to get
-\begin{equation*}
-\norm{H}_{op} \leq \delta s_{n}.
-\end{equation*}
-
-To lower bound the denominator, we note
-\begin{align}
-\abs{\overline{s}_j - s_k} & \geq \abs{\overline{s}_j - \overline{s}_k} - \abs{\overline{s}_k - s_k} \nonumber \\
-& \geq \abs{\overline{s}_j - \overline{s}_k} - \delta \overline{s}_k. \label{eqn:incoherence_proof_8}
-\end{align}
-The eigenvalues of the chain graph are well known to be
-\begin{equation*}
-\overline{s}_j = 2\left(1 - \cos\left(\frac{j\pi}{n}\right)\right) = 4\sin^2\left(\frac{k\pi}{n}\right) ~~\textrm{for $k = 0,\ldots,n - 1$.}
-\end{equation*}
-By Taylor expansion, for $k \leq j \leq n/(2\pi)$ we have
-\begin{equation}
-\label{eqn:incoherence_proof_5}
-\overline{s}_j - \overline{s}_k = 2\left(\cos\left(\frac{k\pi}{n}\right) - \cos\left(\frac{j\pi}{n}\right)\right) \geq \frac{(k - j)^2\pi^2}{n^2}.
-\end{equation}
-and when $j \geq n/(2\pi)$ we have $\overline{s}_j - \overline{s}_k \geq 2(\cos\left(\frac{\pi}{8}\right) - \cos\left(\frac{1}{2}\right) > .09$.
-Additionally since $\sin(x) \leq x$ for all $x \geq 0$, we have
-\begin{equation}
-\label{eqn:incoherence_proof_6}
-\overline{s}_k \leq 4\frac{k^2\pi^2}{n^2}.
-\end{equation}
-Combining~\eqref{eqn:incoherence_proof_5},\eqref{eqn:incoherence_proof_6}, and the lower bound $\abs{k - j} \geq R \geq 8k\sqrt{\delta}$, we have that
-\begin{equation}
-\label{eqn:incoherence_proof_7}
-\abs{\overline{s}_j - \overline{s}_k} - \delta \overline{s}_k \geq .04\frac{(k - j)^2\pi^2}{n^2} - 4 \delta\frac{k^2\pi^2}{n^2} \geq .02\frac{(k - j)^2\pi^2}{n^2}.
-\end{equation}
-The result follows from \eqref{eqn:incoherence_proof_7}, \eqref{eqn:incoherence_proof_8} and \eqref{eqn:davis_kahan}.
-
-\paragraph{Step 3: $L_1$ norm to $L_2$ norm.}
-
-Let $u \in \Reals^n$, and suppose we know that the $L_2$ norm of $u$ is bounded,
-\begin{equation}
-\label{eqn:incoherence_proof_3}
-\norm{u}_2^2 \leq 1.
-\end{equation}
-Under no other conditions on $u$, the upper bound $\norm{u}_1 \leq \sqrt{n}$ is the best that can be hoped for (achieved when $u = (n^{-1/2},\ldots,n^{-1/2})$).
-However, suppose we also know that for some $B_2^2 \geq B_3^2 \geq \ldots \geq B_n^2$, we have that there exists some $R \in [n]$ such that
-\begin{equation}
-\label{eqn:incoherence_proof_4}
-\sum_{j = r}^{n} v_j^2 \leq B_r^2 ~~\textrm{for each $r = R,\ldots,n$.}
-\end{equation}
-Clearly, if $B_n^2 \leq \frac{1}{n}$ then $u$ cannot be equal to $(n^{-1/2},\ldots,n^{-1/2})$.
-We might hope for more general improvements on the bound $\norm{u}_1 \leq \sqrt{n}$ if $B_R^2$ is quite small once $R$ gets sufficiently large.
-The following Lemma gives such a result.
-\begin{lemma}
-	\label{lem:pi_max_pf_1_util_1}
-	Suppose $u \in \Reals^n$ satisfies \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4} for some $1 = B_1^2 = B_2^2 \geq \ldots \geq B_n^2 \geq B_{n+1}^2 = 0$.
-Assume additionally that there exists an $R \in [n]$ such that
-	\begin{equation}
-	\label{eqn:incoherence_util_1}
-	B_r^2 - B_{r + 1}^2 \leq \frac{1}{r}~\textrm{for each $r = R, R+1,\ldots, n - 1$.}
-	\end{equation}
-	Then,
-	\begin{equation}
-	\label{eqn:incoherence_util_2}
-	\norm{u}_1 \leq \sqrt{R}\sqrt{1 - B_R^2} + \sum_{j = R}^{n} \sqrt{B_j^2 - B_{j+1}^2}.
-	\end{equation}
-\end{lemma}
-\begin{proof}
-	Set
-	\begin{equation*}
-	(u_\star)_j^2 = 
-	\begin{cases*}
-	\frac{1}{R}(1 - B_R^2), ~ j \leq R \\
-	B_j^2 - B_{j + 1}^2, j > R
-	\end{cases*}
-	\end{equation*}
-	so that $\norm{u_\star}_1$ is equal to the right hand side of~\eqref{eqn:incoherence_util_2}.
-We now prove by contradiction that $u_\star$ maximizes $\norm{\cdot}_1$ under the constraints \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4}.
-Suppose this were not true, and that the maximum is achieved for some $u \neq u_{\star}$.
-	\begin{itemize}
-		\item since $\norm{u}_1 > \norm{u_{\star}}$ then there exists some index $j$ such that
-		\begin{equation}
-		\label{eqn:incoherence_util_3}
-		u_j > (u_{\star})_j
-		\end{equation} 
-		\item by~\eqref{eqn:incoherence_proof_3} $\norm{u}_2 \leq 1$. Since $\norm{u_{\star}}_2 = 1$, by \eqref{eqn:incoherence_util_3} there must exist some index $k$ such that
-		\begin{equation}
-		\label{eqn:incoherence_util_4}
-		(u_{\star})_k > u_k.
-		\end{equation} 
-		Choose $k$ to be the largest of all such indices.
-		\item suppose $k < j$ and $j > R$.
-Since $k$ was chosen to be the largest such index which satisfied~\eqref{eqn:incoherence_util_4}, clearly $v_i \geq (u_{\star})_i$ for all $i > j$, and by \eqref{eqn:incoherence_util_3} $u_j > (u_{\star})_j$.
-But then
-		\begin{equation*}
-		B_j^2 = \sum_{i = j}^{N} (u_{\star})_i^2 < \sum_{i = j}^{N} u_i^2 
-		\end{equation*}
-		and so $v$ does not satisfy~\eqref{eqn:incoherence_proof_4}. Therefore either $k > j$ or $j \leq R$.
-		\item In either case, we have
-		\begin{equation*}
-		u_j > (u_{\star})_j \geq (u_{\star})_k > u_k.
-		\end{equation*}
-		Let $j \leq l < k$ be the largest index for which $v_j > (a_{\star})_j$.
-
-		We now construct a vector $\wt{u}$ which satisfies the constraints \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4} such that $\norm{\wt{u}}_1 > \norm{u}_1$.
-Once we have shown this, we will have established a contradiction, and the proof of Lemma~\ref{lem:pi_max_pf_1_util_1} will be complete.
-Let $\wt{u} = (\wt{u}_i)$ be given as follows:
-		\begin{equation*}
-		\wt{u}_i = 
-		\begin{cases*}
-		u_i, ~~\textrm{if $i \neq k,l$}, \\
-		u_{\star,k}, ~~\textrm{if $i = k$}, \\
-		\sqrt{u_l^2 - (u_{\star,k}^2 - a_k^2)}, ~~\textrm{if $i = l$.}
-		\end{cases*}
-		\end{equation*}
-		Clearly $\norm{\wt{u}}_2 = \norm{u}_2 = 1$ and so $\wt{u}$ satisfies~\eqref{eqn:incoherence_proof_3}.
-To see that $\wt{u}$ satisfies~\eqref{eqn:incoherence_proof_4} we separate the analysis into cases.
-When $r \geq l + 1$, we have that
-		\begin{equation*}
-		\sum_{i = r}^{n} \wt{u}_i^2 = \sum_{i = r}^{n} u_i^2 \leq B_r^2.
-		\end{equation*}
-		When $l < r \leq k$, we have that
-		\begin{equation*}
-		\sum_{i = r}^{n} \wt{u}_i^2 \leq \sum_{i = r}^{k} u_{\star,i}^2 + \sum_{i = k +1}^{n} u_{i}^2 \leq B_r^2 - B_{k + 1}^2 + B_{k + 1}^2 = B_r^2
-		\end{equation*}
-		Finally, when $R \leq r \leq l$, we have
-		\begin{equation*}
-		\sum_{i = r}^{n} \wt{u}_i^2 \leq u_l^2 - (u_{\star,k}^2 - u_k^2) + (u_{\star,k}^2) + \sum_{i \neq k,l} u_i^2 = \sum_{i = r}^{n} u_i^2 \leq B_r^2.
-		\end{equation*}
-		Therefore $\wt{u}$ satisfies~\eqref{eqn:incoherence_proof_4}. Finally, we have that
-		\begin{equation*}
-		\norm{\wt{u}}_1 = \norm{u}_1 + (u_{\star,k}  - u_k) - (\sqrt{u_l^2 - (u_{\star,k}^2 - u_k^2)} - u_l) > \norm{u}_1 + (u_{\star,k}  - u_k) - (u_{\star,k}  - u_k) = \norm{u}_1,
-		\end{equation*}
-		so we have established the desired contradiction.
-	\end{itemize}
-\end{proof}
-
-\paragraph{Putting the pieces together.}
-We apply Lemma~\ref{lem:pi_max_pf_1_util_1} to the vector $u = (u_i)_{i = 1}^{n}$, where
-\begin{equation*}
-u_i = \frac{1}{\sqrt{2}}\left(\abs{\dotp{v_k}{\overline{v}_{k - i}}} + \abs{\dotp{v_k}{\overline{v}_{k + i}}}\right).
-\end{equation*}
-Note the following:
-\begin{itemize}
-	\item $\norm{u}_2 \leq 1$.
-	\item By \eqref{eqn:incoherence_proof_2}, the vector $u$ satisfies the constraint~\eqref{eqn:incoherence_proof_4} with
-	\begin{equation*}
-	B_r^2 = 400\frac{n^4\delta^2s_{n}^2}{r^4} ~~\textrm{when $r \geq 8 k \sqrt{\delta}$.}
-	\end{equation*}
-	\item Taylor expanding $f(r + 1) = (r + 1)^{-4}$ around $r$, we have that
-	\begin{equation*}
-	B_r^2 - B_{r + 1}^2 \leq 3200 \frac{n^4 \delta^2 s_{n}^2}{r^5} ~~\textrm{when $r \geq 2$.}
-	\end{equation*}
-	Therefore, the sequence $(B_r)$ satisfies \eqref{eqn:incoherence_util_1} for all $r$ large enough such that 
-	\begin{equation*}
-	r \geq 2^{5/4} 100^{1/4} n \sqrt{\delta s_{n}}
-	\end{equation*}
-	\item Since $u$ and $(B_r)$ satisfy the constraints \eqref{eqn:incoherence_proof_3}, \eqref{eqn:incoherence_proof_4} and \eqref{eqn:incoherence_proof_5}, we may apply Lemma~\ref{lem:pi_max_pf_1_util_1} to $u$, obtaining that
-	\begin{equation*}
-	\sum_{j = 1}^{n} \abs{\dotp{v_j}{\overline{v}_k}} = \sqrt{2}\norm{u}_1 \leq \sqrt{2R} + 60\sqrt{2} n^2 \delta s_{n} \sum_{r = R}^{n} r^{-5/2}
-	\end{equation*}
-	for any $R \geq 2 \vee k\sqrt{\delta} \vee 2^{5/4} 100^{1/4} n \sqrt{\delta s_{n}}$. Bounding sum by integral, we have that if $R \geq 2$ then
-	\begin{equation*}
-	\sum_{r = R}^{n} r^{-5/2} \leq \int_{R - 1}^{n - 1} x^{-5/2} \,dx \leq \frac{2^{3/2}2}{3} R^{-3/2}.
-	\end{equation*}
-	Plugging this into the previous expression, we arrive at
-	\begin{equation*}
-	\sum_{j = 1}^{n} \abs{\dotp{\overline{v}_j}{v_k}} \leq \sqrt{R} + 60 \sqrt{2} n^2 \delta s_{\max} \frac{2^{3/2}2}{3R^{3/2}}
-	\end{equation*}
-	for any $R \geq 2 \vee k\sqrt{\delta} \vee 2^{5/4} 100^{1/4} n \sqrt{\delta s_{\max}}$. Lemma~\ref{lem:pi_max_pf_1} then follows from~\eqref{eqn:incoherence_proof_1}.
-\end{itemize}
-
-\section{Auxiliary Results}
-
-\subsection{Results of Others}
-
-Let $\overline{X}$ denote the $n$ evenly spaced grid points on $[0,1]^d$; formally, letting $t = n^{1/d}$
-\begin{equation*}
-\overline{X} = \biggl\{\frac{1}{t}(k_1,\ldots,k_d): k \in [t]^d\biggr\}
-\end{equation*}
-
-\begin{theorem}[Theorem 1 of \textcolor{red}{Garcia-Trillos and Slepcev}]
-	\label{thm:slepcev_transport_distance}
-	With probability at least $1 - c/n$ there exists a bijection $\pi: \overline{X} \to X$ such that
-	\begin{equation}
-	\label{eqn:slepcev_transport_distance}
-	\max_{k \in [t]^d} \abs{\overline{x}_k - \pi(\overline{x}_k)} \leq c \left(\frac{\log n}{n}\right)^{1/d}
-	\end{equation}
-\end{theorem}
-
-\begin{theorem}[\textcolor{red}{Evans} Chapter 5.4, Theorem 1]
-	\label{thm:evans_extension}
-	Assume $U$ is bounded and $\partial U$ is $C^1$.
-Select a bounded open set $V$ such that $U \subset \subset V$ ($U$ is compactly contained in $V$).
-Then there exists a bounded linear operator $E: W^{1,2}(U) \to W^{1,2}(\Rd)$ such that for each $u \in W^{1,2}(U)$:
-	\begin{enumerate}
-		\item $Eu = u$ a.e. in $U$,
-		\item $Eu$ has support within $V$, and 
-		\item 
-		\begin{equation*}
-		\norm{Eu}_{W^{1,2}(\Rd)} \leq C \norm{u}_{W^{1,2}(\Rd)}
-		\end{equation*}
-		the constant $C$ depending only on $U$ and $V$.
-	\end{enumerate}
-\end{theorem}
-
-\begin{lemma}[Poincare inequality for path graphs.]
-	\label{lem: path_poincare}
-	Fix $m \geq 0$. For vertices $V = \set{1, \ldots,m}$ define the path $P(1 \to m) = ((1,2),(2,3),\ldots, (m-1,m))$ and $G_{(1,m)}$ to be the graph consisting only of an edge between $1$ and $m$. Then,
-	\begin{equation*}
-	(m - 1) \cdot P(1 \to m) \succeq G_{(1,m)}
-	\end{equation*}
 \end{lemma}

### 2020-01-30 notes/weekly_notes/1.30.20.NotesForWeek.tex
added higher-order testing rate overly compactly supported Sobolev functions

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{\mathcal{L}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 1/23/20 - 1/30/20}
\author{Alden Green}
\date{\today}
\maketitle

Let $t$ be an integer, $n = t^d$, and 
\begin{equation*}
\overline{X} := \Bigl\{\frac{1}{t}(j_1,\ldots,j_d): k \in [t]^d\Bigr\}
\end{equation*}
consist of $n$ total evenly spaced grid points on $[0,1]^d$. The lattice graph formed over these grid points is 
\begin{equation*}
\overline{G} := \bigl(\overline{X},\overline{E}\bigr),~~ \overline{E} := \set{(\overline{x},\overline{x}'): \norm{\overline{x} - \overline{x}'}_1 = \frac{1}{t}}.
\end{equation*}
We observe $n$ samples according to regression model
\begin{equation}
\label{eqn:grid_regression_model}
y_j = f(\ol{x}_j) + \varepsilon_j, ~\varepsilon_j \overset{\mathrm{i.i.d}}{\sim} \mathcal{N}(0,1) ~~\textrm{for each $j \in [t]^d$,}
\end{equation}
and our goal is to perform the goodness-of-fit test
\begin{equation*}
\mathbb{H}_0 : f = f_0 := 0 ~~\textrm{vs.}~~ \mathbb{H}_a: f \neq f_0.
\end{equation*}
For a given test $\phi:\Reals^n \to \{0,1\}$ and function class $\mathcal{F}$, we evaluate $\phi$ through the lens of worst-case risk:
\begin{equation*}
\mathcal{R}(\phi; \mathcal{F}) := \Ebb_{f_0}(\phi) + \sup_{f \neq f_0 \in \mathcal{F}}\Ebb_{f}(1 - \phi)
\end{equation*}
Unless there exists some separation between functions $f \in \mathcal{F}$ and $f_0$, the type II error can be quite close to $1$. We enforce this separation by removing a ball of radius $\epsilon$ centered at $f_0$ from the function class $\mathcal{F}$, and let
\begin{equation*}
\mathcal{F}_{\epsilon} := \set{f \in \mathcal{F}: \norm{f}_{\Leb^2} \geq \epsilon}.
\end{equation*}
A classical way to study the properties of $\phi$ over $\mathcal{F}$ is through the critical radius: the smallest value of $\epsilon$ such that $\phi$ has small worst-case risk over the remaining function class $\mathcal{R}(\phi;\mathcal{F}_{\epsilon})$, which we shall write as $\mathcal{R}_{\epsilon}(\phi;\mathcal{F})$.

\section{Graph Laplacian Eigenvector Projection Test}
For an arbitrary graph $G$ on $n$ nodes with graph Laplacian $L_G = D_G - A_G$, let $\Lambda(G)$ consist of the eigenvalues $0 = \lambda_1(G) \leq \lambda_2(G) \leq \cdots \leq \lambda_n(G)$ of $L_G$, with $v_k(G) = (v_{k,1}(G),\ldots,v_{k,n}(G)) \in \Reals^n$ denoting the eigenvector corresponding to the $i$th eigenvalues $\lambda_k(G)$. The graph Laplacian eigenvector projection test is a truncated-series test; letting $\kappa$ be some integer between $1$ and $n$ which indicates where we truncate the series $\bigl(Y^T v_k(G)\bigr)$, our test statistic is
\begin{equation*}
T_{\mathrm{spec}}(G) := \frac{1}{n} \sum_{k = 1}^{\kappa} \Biggl(\sum_{i = 1}^{n} y_i v_{k,i}(G)\Biggr)^2
\end{equation*}
and the test $\phi_{\spec}(G) := \1\{T_{\spec}(G) \geq \tau\}$ rejects the null hypothesis when the test statistic $T_{\mathrm{spec}}(G)$ is greater than a pre-specified cutoff $\tau$.

The truncation level $\kappa$ and cutoff $\tau$ are user-specified hyperparameters. When they are appropriately chosen, the test $\phi_{\spec}(\overline{G})$ is minimax optimal over the Holder ball $C^{1}(\Xset;R)$ when $d \leq 4$.
\begin{theorem}
	\label{thm:holder_testing_rate_grid}
	Let $b \geq 1$ and $R > 0$ be fixed constants, and let $d \leq 4$. Suppose we observe data according to the model~\eqref{eqn:grid_regression_model}. Then when the test $\phi_{\spec}(\overline{G})$ is performed with the parameter choices
	\begin{equation*}
	\kappa = n^{2d/(4 + d)},~~ \tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
	\end{equation*}
	the following statement is true: there exists a constant $c$ such that for 
	\begin{equation}
	\label{eqn:holder_testing_rate_grid_1}
	\epsilon^2 \geq c R^2 b^2 n^{-4/(4 + d)}
	\end{equation}
	the worst-case risk of $\phi_{\spec}(\overline{G})$ on $C^{1}(\Xset;R)$ is upper bounded
	\begin{equation}
	\label{eqn:holder_testing_rate_grid_2}
	\mathcal{R}_{\epsilon}\Bigl(\phi_{\spec}(\overline{G});C^{1}(\Xset;R)\Bigr) \leq \frac{1}{b}
	\end{equation}
\end{theorem}
The minimax rate for this problem is established by \textcolor{red}{(Abramovich, De Feis, Sapatinas)}, who consider the more general question of testing over Besov spaces. The test they use to establish the upper bound is related to our own, however, they use a continuum wavelet basis orthogonalized in $L^2([0,1]^d)$ rather than the grid Laplacian eigenvectors we consider.

\section{Proofs}

\subsection{Proof of Theorem~\ref{thm:holder_testing_rate_grid}}

In Lemma~\ref{lem:fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}(G)$ over an arbitrary graph $G$ on $n$ vertices, where we observe data sampled according to the model
\begin{equation}
\label{eqn:regression_model_fixed_graph}
y_i = \beta_i + \varepsilon_i,~~\varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
\end{equation}
Our bound on the Type II error will be stated as a function of $\beta^T L^s \beta$--a measure of the smoothness the signal $\beta$ displays over the graph $G$--as well as the $\kappa$th eigenvalue of the Laplacian $\lambda_{\kappa}(G)$.

\begin{lemma}
	\label{lem:fixed_graph_testing}
	Let $b \geq 1$ be a fixed constant. Suppose we observe data $Y$ according to the model~\eqref{eqn:regression_model_fixed_graph}, and perform the test $\phi_{\spec}(G)$ for some $1 \leq \kappa \leq n$ and
	\begin{equation*}
	\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}.
	\end{equation*}
	Then the following statements are true:
	\begin{enumerate}
		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}(G)$ is upper bounded
		\begin{equation}
		\label{eqn:graph_spectral_type_I_error}
		\mathbb{E}_{\beta_0}\bigl(\phi_{\spec}(G)\bigr) \leq \frac{1}{b^2}.
		\end{equation}
		\item \textbf{Type II error:} For any $\beta \in \Reals^n$ such that
		\begin{equation}
		\label{eqn:fixed_graph_testing_critical_radius}
		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{\beta^T \bigl(L_G\bigr)^s \beta}{n\bigl[\lambda_{\kappa}(G)\bigr]^s}
		\end{equation}
		the Type II error of $\phi_{\spec}$ is upper bounded,
		\begin{equation}
		\label{eqn:graph_spectral_type_II_error}
		\mathbb{E}_{\beta}\bigl(1 - \phi_{\spec}(G)\bigr) \leq \frac{3}{b}.
		\end{equation}
	\end{enumerate}
\end{lemma}
The proof of Lemma~\ref{lem:fixed_graph_testing} can be seen in the \textit{graph\_testing} document. The eigenvalues of the lattice $\ol{G}$ are known exactly, and standard manipulations \textcolor{red}{Sadhanala} lead to the lower bound
\begin{equation}
\label{eqn:eigenvalue_tail_bound_grid}
\lambda_k(\ol{G}) \geq c n^{-2/d}k^{2/d}~\textrm{for $k = 1,\ldots,n$.}
\end{equation}

In light of~\eqref{eqn:eigenvalue_tail_bound_grid} and Lemma~\ref{lem:fixed_graph_testing}, in order to prove Theorem~\ref{thm:holder_testing_rate_grid} it will be sufficient to show the following two inequalities:
\begin{enumerate}[label=(E\arabic*)]
	\item 
	\label{event:test_1}

	For every $f \in C^1(\Xset;R)$,
	\begin{equation}
	\label{eqn:discrete_sobolev_norm_grid}
	f^T \bigl(L_{\ol{G}}\bigr) f \leq 4R^2 n^{1 - 2/d}
	\end{equation}
	\item 
	\label{event:test_2}

	When $d \leq 4$, for every $f \in C^1(\Xset;R)$ satisfying $\norm{f}_{\Leb^2(\Xset)} \geq R^2 n^{-2/(4 + d)}$, additionally
	\begin{equation}
	\label{eqn:l2_to_empirical_norm_grid}
	\frac{1}{n}\sum_{i = 1}^{n} f(x_i)^2 \geq \frac{1}{4}\norm{f}_{\Leb^2(\Xset)}^2
	\end{equation}
\end{enumerate}
Once we have shown these inequalities, some basic algebra yields that for the specific choice $\kappa = (Rn)^{2d/(4s + d)}$ and a sufficiently large choice of $c$ in \eqref{eqn:holder_testing_rate_grid_1},
\begin{align*}
2b\sqrt{\frac{2\kappa}{n^2}} + \frac{f^T \bigl(L_{\ol{G}}\bigr)^s f}{n\bigl[\lambda_{\kappa}(\ol{G})\bigr]^s} \leq \frac{1}{n}\sum_{i = 1}^{n}\bigl[f(x_i)\bigr]^2
\end{align*}
and~\eqref{eqn:holder_testing_rate_grid_2} follows from Lemma~\ref{lem:fixed_graph_testing}.

\paragraph{Proof of~\eqref{eqn:discrete_sobolev_norm_grid}:}
The upper bound~\eqref{eqn:discrete_sobolev_norm_grid} follows straightforwardly from the Holder property,
\begin{align*}
f^T (L_{\ol{G}}) f & = \sum_{i \in [t]^d} \sum_{j \in [t]^d} \bigl(f(\ol{x}_i) - f(\ol{x}_j\bigr)^2 \1\{\norm{i - j}_1 = t\} \\
& \leq R^2 \sum_{i \in [t]^d} \sum_{j \in [t]^d} \norm{\ol{x}_i - \ol{x}_j}_2^2 \cdot  \1\{\norm{i - j}_1 = t\} \\
& = R^2 n^{-2/d} \sum_{i \in [t]^d} \sum_{j \in [t]^d} \1\{\norm{i - j}_1 = t\}  \\
& \leq 4 R^2 n^{1 - 2/d}.
\end{align*}

\paragraph{Proof of~\eqref{eqn:l2_to_empirical_norm_grid}:}
For $x \in \Rd$, let $Q(x)$ be the $d$-dimensional cube of side length $1/t$ with $x$ at its corner,
\begin{equation*}
Q(x) = [x_1 - 1/t,x_1] \otimes \cdots \otimes [x_d - 1/t,x_d]
\end{equation*}
The difference between $f(\ol{x})$ and the average of $f$ over the cube $Q(\ol{x})$ can be bounded using the Holder property,
\begin{equation*}
\int_{Q(\ol{x})} (f(x) - f(\ol{x}))^2 \,dx \leq \frac{1}{n} R^2 n^{-2/d}.
\end{equation*}
Dividing $\Xset$ into cubes and summing over the cubes, we have
\begin{align*}
\norm{f}_{\Leb^2(\Xset)}^2 & = \sum_{j \in [t]^d} \int_{Q(\ol{x}_j)} \bigl[f(x)\bigr]^2 \,dx \\
& \leq 2 \sum_{j \in [t]^d} \frac{1}{n} \Bigl\{\bigl[f(\ol{x}_j)\bigr]^2 + \int_{Q(\ol{x}_j)} \bigl[f(x) - f(\ol{x}_j)]^2 \,dx \Bigr\} \\
& \leq \frac{2}{n}\sum_{j \in [t]^d} \bigl[f(\ol{x}_j)\bigr]^2 + R^2n^{-2/d} \\
& \leq \frac{2}{n}\sum_{j \in [t]^d} \bigl[f(\ol{x}_j)\bigr]^2 + \frac{1}{2}\norm{f}_{\Leb^2(\Xset)}^2
\end{align*}
where the last inequality follows by the assumption $\norm{f}_{\Leb^2(\Xset)}^2 \geq R^2n^{-4/(4 + d)}$ along with the fact that $4/(4 + d) \leq 2/d$ when $d \leq 4$.

\end{document}

### 2020-02-07 notes/graph_testing.tex
added proof of testing rate over W0

@@
 \newcommand{\Qbb}{\mathbb{Q}}
+\newcommand{\Ibb}{\mathbb{I}}
 
@@
 
-In Theorem~\ref{thm:sobolev_testing_rate_order1} we show that under some typical regularity conditions on $P$, the neighborhood graph Laplacian eigenvector projection test $\phi_{\textrm{spec}}(G_{n,r})$ is a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};R)$ for $d = 1,2,3$ and $\Xset = [0,1]^d$.
+In Theorem~\ref{thm:sobolev_testing_rate_order1} we show that under some typical regularity conditions on $P$, the neighborhood graph Laplacian eigenvector projection test $\phi_{\textrm{spec}}(G_{n,r})$ is a minimax optimal test over the Sobolev ball $H^1(\mathcal{X};R)$ for $d = 1,2,3$ and $\Xset = [0,1]^d$.
 Let $p_d$ equal $3/4$ when $d = 2$ and equal $1/d$ for $d \geq 3$.
@@
 	\label{eqn:sobolev_testing_rate_order1_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}(G_{n,r}); W^{1,2}(\mathcal{X};R)) \leq \frac{c_2}{b}.
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}(G_{n,r}); H^1(\mathcal{X};R)) \leq \frac{c_2}{b}.
 	\end{equation}
@@
 The testing problem in the random design case with known variance fundamentally changes in the low-smoothness regime, where the number of bounded derivatives is small compared to the dimension $d$.
-In the case where $f \in W^{1,2}(\Xset;R)$ is assumed to have one derivative which is bounded in $\Leb^2(\Xset)$ norm, this transition occurs at $d = 4$.
+In the case where $f \in H^1(\Xset;R)$ is assumed to have one derivative which is bounded in $\Leb^2(\Xset)$ norm, this transition occurs at $d = 4$.
 We shall discuss testing in the low-smoothness regime shortly.
+First, let's consider the situation when we assume more derivatives worth of smoothness.
 
@@
 
-When we additionally believe $W^{s,2}(\Xset)$, we know (again from Ingster) that the minimax testing rate is $n^{-4s/(4s + d)}$.
-Now, however, our estimates of the eigenvectors no longer track the derivatives finely enough at the boundaries of $\Xset$ to establish the same minimax rate over all $W^{s,2}(\mathcal{X})$.
-However, if we assume $f$ and its derivatives are supported only on the interior of $\mathcal{X}$, we recover the usual rate.
-Theorem~\ref{thm:sobolev_testing_rate} presents our formal result, that $\phi_{\textrm{spec}}(G_{n,r})$ is a minimax optimal test over the Sobolev balls $W_{\sigma}^{s,2}(\mathcal{X};R)$ whenever $4s > d$ and $\Xset = [0,1]^d$.
+When we additionally believe $H^s(\Xset)$, we know (again from Ingster) that the minimax testing rate is $n^{-4s/(4s + d)}$.
+Now, however, our estimates of the eigenvectors no longer track the derivatives finely enough at the boundaries of $\Xset$ to establish the same minimax rate over all $H^s(\mathcal{X})$.
+However, if we assume $f$ and its derivatives are compactly supported on $\Xset$, we recover the usual rate.
+Theorem~\ref{thm:sobolev_testing_rate} presents our formal result, that $\phi_{\textrm{spec}}(G_{n,r})$ is a minimax optimal test over the Sobolev balls $H_0^s(\mathcal{X};R)$ whenever $4s > d$ and $\Xset = [0,1]^d$.
 
@@
 	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}.
-Let $R,\sigma > 0$, $b,s,d \geq 1$ be fixed constants, with $s$ and $d$ integers.
+Let $R > 0$, $b,s,d \geq 1$ be fixed constants, with $s$ and $d$ integers.
 Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p \in C^{s-1}(\Xset;p_{\max})$ for some $p_{\max} < \infty$, and further that $p(x)$ is bounded away from zero, i.e.
@@
 	\begin{equation*}
-	n^{-1/(2(s-1) + d)} \leq r(n) \leq \min\bigl\{n^{-4/((4s + d)(2+d))},\sigma/s\bigr\}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	n^{-1/(2(s-1) + d)} \leq r(n) \leq n^{-4/((4s + d)(2+d))}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
 	\end{equation*}
@@
 	\label{eqn:sobolev_testing_rate_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W_{\sigma}^{s,2}(\mathcal{X};R)) \leq \frac{c_2}{b}.
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; H_0^{s}(\mathcal{X};R)) \leq \frac{c_2}{b}.
 	\end{equation}
@@
 \begin{equation*}
-W_d^{s,2}(\mathcal{X}) \subseteq \mathcal{L}_d^4(\mathcal{X}) 
+H^{s}(\mathcal{X}) \subseteq \Leb^4(\mathcal{X}) 
 \end{equation*}
@@
 	\begin{equation*}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}. 
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; H^1(\mathcal{X};R)) \lesssim \frac{1}{b}. 
 	\end{equation*}
@@
 
-In Theorem~\ref{thm:twosample_sobolev_testing_rate} we show that under some typical regularity conditions on $P$, the test $\phi_{\textrm{spec}}^{(2)} := \1\{T_{\mathrm{spec}}^{(2)} \geq \tau\}$ is, up to log factors, a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};R)$ when $d = 1$.
+In Theorem~\ref{thm:twosample_sobolev_testing_rate} we show that under some typical regularity conditions on $P$, the test $\phi_{\textrm{spec}}^{(2)} := \1\{T_{\mathrm{spec}}^{(2)} \geq \tau\}$ is, up to log factors, a minimax optimal test over the Sobolev ball $H^1(\mathcal{X};R)$ when $d = 1$.
 
@@
 	\item Are either or both of the test statistics \eqref{eqn:graph_spectral_projections_2}-\eqref{eqn:sobolev_IPM} minimax optimal over $W_d^{1,2}(\mathcal{X};R)$ when $d \leq 4$?
-	\item Is the test statistic \eqref{eqn:graph_spectral_projections_2}, computed over a graph with suitable choice of kernel $K$, minimax optimal over higher order derivative classes $C_d^s$ and $W^{s,2}$?
+	\item Is the test statistic \eqref{eqn:graph_spectral_projections_2}, computed over a graph with suitable choice of kernel $K$, minimax optimal over higher order derivative classes $C_d^s$ and $H^s$?
 	\item Modify the test statistic \eqref{eqn:sobolev_IPM} by replacing the function class $\Theta_{1,2}$ with
@@
 	\end{equation}
-	Is the modified test statistic, computed over a graph with suitable choice of kernel $K$, minimax optimal over higher order derivative classes $C_d^s$ and $W^{s,2}$?
+	Is the modified test statistic, computed over a graph with suitable choice of kernel $K$, minimax optimal over higher order derivative classes $C_d^s$ and $H^s$?
 	\item What is the minimax testing rate over $BV_d^{1}(\mathcal{X};R)$? Does it exhibit a phase transition analogous to the minimax estimation rate over bounded variation spaces?
@@
 \begin{equation*}
-\norm{f}_{C_d^{s}(\mathcal{X})} := \sum_{\abs{\alpha} < s} \norm{D^{\alpha}f}_{\infty} + \sum_{\abs{\alpha} = \ell} \sup_{x,y \in \mathcal{X}} \frac{\abs{D^{\alpha}f(y) - D^{\alpha}f(x)}}{\norm{x - y}_2^{s - \ell}}
-\end{equation*}
-and the $s$th Holder space $C_d^{s}(\mathcal{X})$ consists of all functions which are $\ell$ times continuously differentiable with finite Holder norm $\norm{\cdot}_{C_d^{s}(\mathcal{X})}$.
-Denote the Holder ball by $C_d^{s}(\mathcal{X},R) = \set{f \in C_d^{s}(\mathcal{X}): \norm{f}_{C_d^{s}(\mathcal{X})} \leq R}$.
+\norm{f}_{C^{s}(\mathcal{X})} := \sum_{\abs{\alpha} < s} \norm{D^{\alpha}f}_{\infty} + \sum_{\abs{\alpha} = \ell} \sup_{x,y \in \mathcal{X}} \frac{\abs{D^{\alpha}f(y) - D^{\alpha}f(x)}}{\norm{x - y}_2^{s - \ell}}
+\end{equation*}
+and the $s$th Holder space $C^{s}(\mathcal{X})$ consists of all functions which are $\ell$ times continuously differentiable with finite Holder norm $\norm{\cdot}_{C^{s}(\mathcal{X})}$.
+Denote the Holder ball by $C_d^{s}(\mathcal{X},R) = \set{f \in C^{s}(\mathcal{X}): \norm{f}_{C_d^{s}(\mathcal{X})} \leq R}$.
+Let $C_c^{s}(\Xset)$ consist of those functions in $C^s(\Xset)$ which are compactly supported within $\Xset$; formally $f \in C_c^{s}(\Xset)$ if and only if $f \in C^s(\Xset)$ and there exists an open set $V \subset \ol{V} \subset \Xset$ such that $\mathrm{supp}(f) \subset V$.
 
@@
 
-For a given $s > 0$, the Sobolev space $W^{s,2}(\mathcal{X})$ consists of all functions $f \in \Leb^2(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^2(\mathcal{X})$.
-The Sobolev $\{s,2\}$ norm is then
-\begin{equation*}
-\norm{f}_{W^{s,2}(\mathcal{X})}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^2 \,dx
-\end{equation*}
-and for a given $R > 0$, the corresponding ball is $W^{s,2}(\Xset; R) = \set{f: \norm{f}_{W^{s,2}(\Xset)} \leq R}$.
-
-For a given $\sigma > 0$, let $\Xset_{\sigma} := \set{x \in \Xset: \mathrm{dist}(x,\partial\Xset) > \sigma}$.
-Let $W_{\sigma}^{s,2}(\Xset;L)$ consist of all those functions $f \in W^{s,2}(\Xset)$ such that $\mathrm{supp}(f) \subseteq \Xset_{\sigma}$.
+For a given $s$ and $p > 0$, the Sobolev space $W^{s,p}(\mathcal{X})$ consists of all functions $f \in \Leb^2(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^2(\mathcal{X})$.
+The Sobolev $\{s,p\}$ norm is then
+\begin{equation*}
+\norm{f}_{W^{s,p}(\mathcal{X})}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^p \,dx
+\end{equation*}
+and for a given $R > 0$, the corresponding ball is $W^{s,p}(\Xset; R) = \set{f: \norm{f}_{H^s(\Xset)} \leq R}$.
+In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space; we adopt the usual convention of writing $H^s(\Xset) = H^s(\Xset)$.
+
+We write $W_0^{s,p}(\Xset)$ for the completion of $C_c^{\infty}(\Xset)$, and let $H_0(\Xset) = W_0^{s,2}(\Xset)$.
 
@@
 
-We will treat $f$ interchangeably as a function $f:\Rd \to \Reals$, and as a vector $f = (f(x_1),\ldots,f(x_n))$. It should always be clear from context how we are using $f$.
+We will treat $f$ interchangeably as a function $f:\Rd \to \Reals$, and as a vector $f = (f(x_1),\ldots,f(x_n))$.
+More generally, for $U \subset V$ and $f:V \to \Reals$, the action of $f$ on $U$ is naturally defined via the restriction operator $R_U$; we will forego additional notation and simply treat $f$ as a function on $U$, defined as $f(x) = R_Uf(x)$ for $x \in U$.
+It should always be clear from context how we are using $f$.
+
 For sequences $(a_n), (b_n)$, we say $a_n = O(b_n)$ if there exists a constant $c$ such that $a_n \leq cb_n$ for all $c$. We use the notation $a \wedge b$ for the minimum of $a$ and $b$.
@@
 	\label{eqn:continuous_to_discrete_sobolev_norm_order1}
-	f^T L f \leq c \cdot b \cdot \norm{f}_{W^{1,2}(\Xset)}^2 \cdot n^{2} r^{d + 2} 
+	f^T L f \leq c \cdot b \cdot \norm{f}_{H^1(\Xset)}^2 \cdot n^{2} r^{d + 2} 
 	\end{equation}
@@
 \subsubsection{Proof of \eqref{eqn:l2_to_empirical_norm_order1}}
-The inequality~\eqref{eqn:l2_to_empirical_norm_order1} is implied by the more general Lemma~\ref{lem:empirical_norm_sobolev}, which we state in Section~\ref{subsec:sobolev_testing_rate_pf} and use to prove a similar inequality when $f \in W^{s,2}(\Xset)$ for general $s$.
+The inequality~\eqref{eqn:l2_to_empirical_norm_order1} is implied by the more general Lemma~\ref{lem:empirical_norm_sobolev}, which we state in Section~\ref{subsec:sobolev_testing_rate_pf} and use to prove a similar inequality when $f \in H^s(\Xset)$ for general $s$.
 
@@
 	\label{event:discrete_sobolev_norm}
-	\textbf{Graph Sobolev norm:} For any $n^{-1/(2(s - 1) + d)}\leq r \leq \sigma/s$,
+	\textbf{Graph Sobolev norm:} For any $n^{-1/(2(s - 1) + d)}\leq r$ such that $r \to 0$ as $n \to \infty$,
 	\begin{equation}
 	\label{eqn:continuous_to_discrete_sobolev_norm}
-	f^T L^s f \leq c \cdot b \cdot \norm{f}_{W^{s,2}(\Xset)}^2 \cdot n^{s + 1} r^{s(d + 2)} 
-	\end{equation}
+	f^T L^s f \leq c \cdot b \cdot \norm{f}_{H^s(\Xset)}^2 \cdot n^{s + 1} r^{s(d + 2)}.
+	\end{equation}
+	for all $n$ sufficiently large.
 	\item 
@@
 	\label{lem:roughness_functional_expectation_sobolev}
-	Let $\Xset$ be a Lipschitz domain, let $s \geq 1$ be an integer, and let $0 < \sigma < 1$.
-Suppose that $f \in W_{\sigma}^{s,2}(\Xset)$, and further that $p \in C^{s-1}(\Xset;p_{\max})$ for some constant $p_{\max} < \infty$.
-Then for any $2$nd-order kernel $K$ and any $n^{-1/(2(s - 1) + d)} \leq r(n) < \sigma/s$, the expected graph Sobolev seminorm is upper bounded
+	Let $\Xset$ be a Lipschitz domain and let $s \geq 1$ be an integer.
+Suppose that $f \in H_0^{s}(\Xset)$, and further that $p \in C^{s-1}(\Xset;p_{\max})$ for some constant $p_{\max} < \infty$.
+Then for any $2$nd-order kernel $K$ and any $n^{-1/(2(s - 1) + d)} \leq r(n)$ such that $r \to 0$ as $n \to \infty$, the expected graph Sobolev seminorm is upper bounded
 	\begin{equation}
 	\label{eqn:roughness_functional_expectation_sobolev}
-	\Ebb\bigl[f^T L^s f\bigr] \leq c \cdot \norm{f}_{W^{s,2}(\Xset)}^2 \cdot n^{s + 1}r^{s(d + 2)}
-	\end{equation}
+	\Ebb\bigl[f^T L^s f\bigr] \leq c \cdot \norm{f}_{H^s(\Xset)}^2 \cdot n^{s + 1}r^{s(d + 2)}
+	\end{equation}
+	for all $n$ sufficiently large.
 \end{lemma}
 
-We note that the proof of Lemma~\ref{lem:roughness_functional_expectation_sobolev} is where we rely on the fact that $f$ and its derivatives are supported on $\Xset_{\sigma}$.
+We note that the proof of Lemma~\ref{lem:roughness_functional_expectation_sobolev} is where we rely on the fact that $f$ and its derivatives are compactly supported on $\Xset$.
 The proof of Lemma~\ref{lem:roughness_functional_expectation_sobolev} is lengthy, and we defer it to Section~\ref{sec:technical_lemma_proofs}.
@@
 	\end{equation*}
-	and let $f \in W^{s,2}(\Xset)$.Then for any $b \geq 1$, there exists $c_1$ such that if 
+	and let $f \in H^s(\Xset)$.Then for any $b \geq 1$, there exists $c_1$ such that if 
 	\begin{equation}
@@
 	\begin{cases*}
-	c_1 \cdot b \cdot \norm{f}_{W^{s,2}(\Xset)} \cdot \max\{n^{-1/2},n^{-s/d}\},~~\textrm{if}~2s \neq d \\
-	c_1 \cdot b \cdot \norm{f}_{W^{s,2}(\Xset)} \cdot n^{-a/2},~~\textrm{if}~ 2s = d ~\textrm{for any}~ 0 < a < 1
+	c_1 \cdot b \cdot \norm{f}_{H^s(\Xset)} \cdot \max\{n^{-1/2},n^{-s/d}\},~~\textrm{if}~2s \neq d \\
+	c_1 \cdot b \cdot \norm{f}_{H^s(\Xset)} \cdot n^{-a/2},~~\textrm{if}~ 2s = d ~\textrm{for any}~ 0 < a < 1
 	\end{cases*}
@@
 \begin{equation*}
-\sup_{x \in \Xset} \abs{f(x)} \leq \norm{f}_{C^{\gamma}(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}.
+\sup_{x \in \Xset} \abs{f(x)} \leq \norm{f}_{C^{\gamma}(\Xset)} \leq c \norm{f}_{H^s(\Xset)}.
 \end{equation*}
@@
 & \leq \left(\sup_{x \in \Xset} \abs{f(x)}\right)^2 \cdot \int_{\Xset} [f(x)]^2 \,dx \\
-& \leq c \norm{f}_{W^{s,2}(\Xset)}^2 \cdot \norm{f}_{\Leb^2(\Xset)}^2.
+& \leq c \norm{f}_{H^s(\Xset)}^2 \cdot \norm{f}_{\Leb^2(\Xset)}^2.
 \end{align*}
@@
 \begin{equation*}
-p_{\max}^2\frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c \norm{f}_{W^{s,2}(\Xset)}^2 \cdot \frac{\norm{f}_{\Leb^2(\Xset)}^4}{n \norm{f}_{\Leb^2(\Xset)}^2} \leq c \frac{\norm{f}_{\Leb^2(\Xset)}^4}{c_1^2 b^2} \leq \frac{\Ebb\bigl[\norm{f}_n^2\bigr]}{b^2},
+p_{\max}^2\frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c \norm{f}_{H^s(\Xset)}^2 \cdot \frac{\norm{f}_{\Leb^2(\Xset)}^4}{n \norm{f}_{\Leb^2(\Xset)}^2} \leq c \frac{\norm{f}_{\Leb^2(\Xset)}^4}{c_1^2 b^2} \leq \frac{\Ebb\bigl[\norm{f}_n^2\bigr]}{b^2},
 \end{equation*}
@@
 \begin{equation*}
-\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}.
+\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{H^s(\Xset)}.
 \end{equation*}
@@
 \begin{equation*}
-\norm{f}_{\Leb^4(\Xset)}^4 \leq \norm{f}_{\Leb^2}^{2\theta} \cdot \norm{f}_{\Leb^q(\Xset)}^{(1 - \theta)q} \leq c \norm{f}_{\Leb^2(\Xset)}^{4} \cdot \left(\frac{\norm{f}_{W^{s,2}(\Xset)}}{\norm{f}_{\Leb^2(\Xset)}}\right)^{d/s}.
-\end{equation*}
-By assumption, $\norm{f}_{\Leb^2(\Xset)} \geq c_1 b \norm{f}_{W^{s,2}(\Xset)} n^{-s/d}$, and therefore
-\begin{equation*}
-p_{\max}^2 \frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c\norm{f}_{\Leb^2(\Xset)}^4 \left(\frac{\norm{f}_{W^{s,2}(\Xset)}}{n^{s/d}\norm{f}_{\Leb^2(\Xset)}}\right)^{d/s} \leq \frac{c\norm{f}_{\Leb^2(\Xset)}^4}{c_1b^{d/s}} \leq \frac{\norm{f}_{\Leb^2(\Xset)}^4}{b^2}.
+\norm{f}_{\Leb^4(\Xset)}^4 \leq \norm{f}_{\Leb^2}^{2\theta} \cdot \norm{f}_{\Leb^q(\Xset)}^{(1 - \theta)q} \leq c \norm{f}_{\Leb^2(\Xset)}^{4} \cdot \left(\frac{\norm{f}_{H^s(\Xset)}}{\norm{f}_{\Leb^2(\Xset)}}\right)^{d/s}.
+\end{equation*}
+By assumption, $\norm{f}_{\Leb^2(\Xset)} \geq c_1 b \norm{f}_{H^s(\Xset)} n^{-s/d}$, and therefore
+\begin{equation*}
+p_{\max}^2 \frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c\norm{f}_{\Leb^2(\Xset)}^4 \left(\frac{\norm{f}_{H^s(\Xset)}}{n^{s/d}\norm{f}_{\Leb^2(\Xset)}}\right)^{d/s} \leq \frac{c\norm{f}_{\Leb^2(\Xset)}^4}{c_1b^{d/s}} \leq \frac{\norm{f}_{\Leb^2(\Xset)}^4}{b^2}.
 \end{equation*}
@@
 
-Now, we bound Type II error under the assumption $f \in W_{\sigma}^{s,2}(\mathcal{X};R), p \in C^{s - 1}(\Xset;p_{\max})$ uniformly bounded away from $0$ and $\infty$ over $\Xset$, and 
+Now, we bound Type II error under the assumption $f \in H_0^{s}(\mathcal{X};R), p \in C^{s - 1}(\Xset;p_{\max})$ uniformly bounded away from $0$ and $\infty$ over $\Xset$, and 
 \begin{equation}
@@
 \end{equation}
-Choosing $n^{-1/(2(s - 1) + d)}\leq r(n) \leq n^{-4/((2+d)(4s + d))}$, we may therefore apply our conclusions in Step 2; namely, that for every possible choice of $f \in W^{s,2}(\Xset;R)$ there exists a good set $\mathcal{E}_f \subseteq \Xset^n$ with $\Pbb(\mathcal{E}_f) \geq 1 - c/b$ such that each of \eqref{eqn:continuous_to_discrete_sobolev_norm}, \eqref{eqn:eigenvalue_tail_bound}, and \eqref{eqn:l2_to_empirical_norm_sobolev} hold for all $X \subseteq \mathcal{E}_f$.
+Choosing $n^{-1/(2(s - 1) + d)}\leq r(n) \leq n^{-4/((2+d)(4s + d))}$, we may therefore apply our conclusions in Step 2; namely, that for every possible choice of $f \in H^s(\Xset;R)$ there exists a good set $\mathcal{E}_f \subseteq \Xset^n$ with $\Pbb(\mathcal{E}_f) \geq 1 - c/b$ such that each of \eqref{eqn:continuous_to_discrete_sobolev_norm}, \eqref{eqn:eigenvalue_tail_bound}, and \eqref{eqn:l2_to_empirical_norm_sobolev} hold for all $X \subseteq \mathcal{E}_f$.
 The choice $\kappa = (nR^2)^{2d/(4s + d)}$ balances the squared bias and variance terms on the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius}, and we have that for all $X \subseteq \mathcal{E}_f$
@@
 \begin{equation*}
-\sup_{\substack{f \in W^{s,2}(\Xset;R), \\ \norm{f}_{\Leb^2(\Xset)} \geq \epsilon}}\mathbb{E}_{f}(1 - \phi_{\spec}) \leq \sup_{\substack{f \in W^{s,2}(\Xset;R), \\ \norm{f}_{\Leb^2(\Xset)} \geq \epsilon}} \mathbb{E}\bigl[\mathbb{E}_{\beta}(1 - \phi_{\spec}|X \in \mathcal{E}_f)\bigr] + \frac{c}{b} \leq \frac{3 + c}{b},
+\sup_{\substack{f \in H^s(\Xset;R), \\ \norm{f}_{\Leb^2(\Xset)} \geq \epsilon}}\mathbb{E}_{f}(1 - \phi_{\spec}) \leq \sup_{\substack{f \in H^s(\Xset;R), \\ \norm{f}_{\Leb^2(\Xset)} \geq \epsilon}} \mathbb{E}\bigl[\mathbb{E}_{\beta}(1 - \phi_{\spec}|X \in \mathcal{E}_f)\bigr] + \frac{c}{b} \leq \frac{3 + c}{b},
 \end{equation*}
@@
 
-Note that since $p,q \in W^{1,2}(\mathcal{X};R)$, we have $f := p - q \in W^{1,2}(\mathcal{X};2R)$.
+Note that since $p,q \in H^1(\mathcal{X};R)$, we have $f := p - q \in H^1(\mathcal{X};2R)$.
 Therefore the bounds~\eqref{eqn:continuous_to_discrete_sobolev_norm} and \eqref{eqn:l2_to_empirical_norm} apply to $S_2(\varDelta,G_{n,r})$ and $(1/n)\sum_{i = 1}^{n}\varDelta_i^2$, respectively.
@@
 \begin{cases*}
-O(r^{2s}) \cdot \norm{f}_{W^{s,2}(\Xset)}^2, & ~~\textrm{if $I = s + 1$} \\
-O(r^{2} r^{d(I - (s + 1))}) \cdot [f]_{W^{1,2}(\Xset)}^2, & ~~\textrm{if $I < s + 1$}~ 
+O(r^{2s}) \cdot \norm{f}_{H^s(\Xset)}^2, & ~~\textrm{if $I = s + 1$} \\
+O(r^{2} r^{d(I - (s + 1))}) \cdot [f]_{H^1(\Xset)}^2, & ~~\textrm{if $I < s + 1$}~ 
 \end{cases*}
@@
 In the sum on the right hand side of~\eqref{eqn:continuous_to_discrete_sobolev_norm_pf1}, there are $O(n^{I})$ terms with exactly $I$ distinct indices.
-When $I < s + 1$, by~\eqref{eqn:expected_difference_operators_1} the total contribution of such terms to the sum is $O(n^{I}r^{d(I - 1) + 2}) \cdot [f]_{W^{1,2}(\Xset)}^2$.
+When $I < s + 1$, by~\eqref{eqn:expected_difference_operators_1} the total contribution of such terms to the sum is $O(n^{I}r^{d(I - 1) + 2}) \cdot [f]_{H^1(\Xset)}^2$.
 Since by assumption $r \geq n^{-1/d}$, this increases with $I$.
-Taking $I = s$ to be the largest integer less than $s + 1$, the contribution of these terms to the sum is therefore $O(n^sr^{d(s - 1) + 2}) \cdot [f]_{W^{1,2}(\Xset)}^2$ which in light of the restriction $r \geq n^{-1/(2(s - 1) + d)}$ is $O(n^{s+1}r^{s(d +2)}) \cdot [f]_{W^{1,2}(\Xset)}^2$.
-On the other hand when $I = s + 1$, by~\eqref{eqn:expected_difference_operators_1} we immediately have that the total contribution of these terms is $O(n^{s + 1}r^{2(s + d)}) \cdot \norm{f}_{W^{s,2}(\Xset)}$.
+Taking $I = s$ to be the largest integer less than $s + 1$, the contribution of these terms to the sum is therefore $O(n^sr^{d(s - 1) + 2}) \cdot [f]_{H^1(\Xset)}^2$ which in light of the restriction $r \geq n^{-1/(2(s - 1) + d)}$ is $O(n^{s+1}r^{s(d +2)}) \cdot [f]_{H^1(\Xset)}^2$.
+On the other hand when $I = s + 1$, by~\eqref{eqn:expected_difference_operators_1} we immediately have that the total contribution of these terms is $O(n^{s + 1}r^{2(s + d)}) \cdot \norm{f}_{H^s(\Xset)}$.
 Therefore,
 \begin{equation*}
-\Ebb(f^T L^s f) = O(n^{s+1}r^{s(d+2)}) \cdot \norm{f}_{W^{s,2}(\Xset)}.
+\Ebb(f^T L^s f) = O(n^{s+1}r^{s(d+2)}) \cdot \norm{f}_{H^s(\Xset)}^2.
 \end{equation*}
@@
 \begin{cases*}
-O(r^{2s}) \cdot \norm{f}_{W^{s,2}(\Xset)}^2, & ~~\textrm{if $I = s + 1$} \\
-O(r^{2} \cdot r^{d(I - (s + 1))}) \cdot [f]_{W^{1,2}(\Xset)}^2, & ~~\textrm{if $I < s + 1$}~ 
+O(r^{2s}) \cdot \norm{f}_{H^s(\Xset)}^2, & ~~\textrm{if $I = s + 1$} \\
+O(r^{2} \cdot r^{d(I - (s + 1))}) \cdot [f]_{H^1(\Xset)}^2, & ~~\textrm{if $I < s + 1$}~ 
 \end{cases*}
@@
 
-Note that if $f$ is constant almost everywhere in $\Xset$, the claim is immediate as $D_kf(x_i) = 0$ with probability one. Otherwise $[f]_{W^{1,2}(\Xset)} > 0$, which we shall assume in what follows.
-
-Let $\delta = \min\{r^{2s},1\}\cdot[f]_{W^{1,2}(\Xset)} > 0$.
-Our analysis will make heavy use of Taylor expansions, and we therefore would like to show that there exists some $g \in C^{s}(\Xset) \cap W_{\sigma/2}^{s,2}(\Xset)$ such that
-\begin{equation*}
-\Bigl|\Ebb[D_kf(x_i) D_{\ell} f(x_i)] - \Ebb[D_kg(x_i) D_\ell g(x_i)]\Bigr| <  \delta, ~~\textrm{and}~~[g]_{W^{\ell,2}(\Rd)} \leq c [f]_{W^{\ell,2}(\Rd)}~\textrm{for each $\ell = 0,\ldots,s$.}
-\end{equation*}
-Then if \eqref{eqn:expected_difference_operators_1} and \eqref{eqn:expected_difference_operators_2} hold with respect to $g$, they hold (up to constants) with respect to $f$ as well. 
-Note that since $f$ is supported on $\Xset_{\sigma}$, the extension
-\begin{equation*}
-\wt{f} := 
-\begin{cases*}
-f(x),~~ x \in \Xset_{\sigma} \\
-0,~~ x \in \Rd - \Xset_{\sigma}
-\end{cases*}
-\end{equation*}
-satisfies $\wt{f}(x) = f(x)$ for all $x \in \Xset$, and $\norm{\wt{f}}_{W^{s,2}(\Rd)} = \norm{f}_{W^{s,2}}(\Xset)$.
-
-To construct our function $g$, we consider as candidates the convolutions
-\begin{equation}
-g_m := \wt{f} \ast \eta_{1/m},~~\textrm{for each $m \in \mathbb{N}$.}
-\end{equation}
-where $\eta$ is the standard mollifier (for a definition see \textcolor{red}{Evans}.) We make the following observations:
-\begin{itemize}
-	\item Since $\eta$ is a smooth function, $g_m \in C^{\infty}(\Rd)$ and therefore $g_m \in C^s(\Xset)$. 
-	\item For all sufficiently large $m$ and all multiindices $\alpha$,
-	\begin{equation*}
-	g_m^{(\alpha)}(x) = 0~~\textrm{if $x \in \Reals^d - X_{\sigma/2}$}.
-	\end{equation*}
-	and along with the previous observation this implies $g_m \in W_{\sigma/2}^{s,2}(\Xset)$. 
-	\item By Theorem~1 of \textcolor{red}{Evans 5.3.1}, $\norm{g_m - \wt{f}}_{W_{\mathrm{loc}}^{s,2}(\Rd)} \to 0$ as $m \to \infty$, which implies  $\norm{g_m - f}_{W^{s,2}(\Xset)} \to 0$.
-\end{itemize}
-By the Cauchy-Schwarz inequality
-\begin{equation*}
-\abs{\Ebb\bigl[D_kf(x_i) D_{\ell} f(x_i)\bigr] - \Ebb\bigl[D_kg_m(x_i) D_\ell g_m(x_i)\bigr]} \leq \frac{c}{r^{sd}} \cdot \norm{f - g_m}_{\Leb^2(\Xset)}
-\end{equation*}
-and taking $m$ to be sufficiently large, we can make the right hand side less than $\delta$.
-On the other hand, since $\norm{g_m - f}_{W^{s,2}(\Xset)} \to 0$, there exists $m$ sufficiently large such that $[g_m]_{W^{\ell,2}(\Xset)}$ is at most say $2 [f]_{W^{\ell,2}(\Xset)}$ for each $\ell = 0,\ldots,s$.
-We therefore take $g = g_m$ for $m$ large enough to satisfy both conditions.
-
-Our task is now to prove that \eqref{eqn:expected_difference_operators_1} and \eqref{eqn:expected_difference_operators_2} hold with respect to $g$.
 We first prove the desired bounds in the case when some indices are repeated, and then the desired bounds in the case when all indices are distinct.
@@
 When at least one index is repeated, we obtain a sufficient upper bound by reducing the problem of upper bounding the iterated difference operator to that of upper bounding a single difference operator.
-Letting $k = (k_1,\ldots,k_q)$, we can show by induction that the absolute value of the iterated difference operator $\abs{D_kg(x_i)}$ is upper bounded by
-\begin{equation*}
-\abs{D_kg(x_i)} \leq \left(\frac{2K_{\max}}{r^d}\right)^{q-1} \sum_{h \in k \cup i} \abs{D_{k_q}g(x_h)} \cdot \1\{G_{n,r}[X_{k \cup i}]~\textrm{is a connected graph} \}.
+Letting $k = (k_1,\ldots,k_q)$, we can show by induction that the absolute value of the iterated difference operator $\abs{D_kf(x_i)}$ is upper bounded by
+\begin{equation*}
+\abs{D_kf(x_i)} \leq \left(\frac{2K_{\max}}{r^d}\right)^{q-1} \sum_{h \in k \cup i} \abs{D_{k_q}f(x_h)} \cdot \1\{G_{n,r}[X_{k \cup i}]~\textrm{is a connected graph} \}.
 \end{equation*}
@@
 \begin{align}
-\abs{D_kg(x_i)} \cdot \abs{D_{\ell}g(x_i)} & \leq \left(\frac{2K_{\max}}{r^d}\right)^{2(q - 1)} \sum_{h,j \in k \cup \ell \cup i} \abs{D_{k_q}g(x_h)} \cdot \abs{D_{\ell_q}g(x_j)} \cdot \1\{G_{n,r}[X_{k \cup i}], G_{n,r}[X_{\ell \cup i}]~\textrm{are connected graphs.} \} \nonumber \\
-& =  \left(\frac{2K_{\max}}{r^d}\right)^{2(q - 1)} \sum_{h,j \in k \cup \ell \cup i} \abs{D_{k_q}g(x_h)} \cdot \abs{D_{\ell_q}g(x_j)} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is a connected graph.} \} \label{eqn:expected_difference_operators_sobolev_pf0}
+\abs{D_kf(x_i)} \cdot \abs{D_{\ell}f(x_i)} & \leq \left(\frac{2K_{\max}}{r^d}\right)^{2(q - 1)} \sum_{h,j \in k \cup \ell \cup i} \abs{D_{k_q}f(x_h)} \cdot \abs{D_{\ell_q}f(x_j)} \cdot \1\{G_{n,r}[X_{k \cup i}], G_{n,r}[X_{\ell \cup i}]~\textrm{are connected graphs.} \} \nonumber \\
+& =  \left(\frac{2K_{\max}}{r^d}\right)^{2(q - 1)} \sum_{h,j \in k \cup \ell \cup i} \abs{D_{k_q}f(x_h)} \cdot \abs{D_{\ell_q}f(x_j)} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is a connected graph.} \} \label{eqn:expected_difference_operators_sobolev_pf0}
 \end{align}
@@
 \begin{equation*}
-\Ebb\Bigl[\abs{D_{k_q}g(x_h)} \cdot \abs{D_{\ell_q}g(x_j)}\Bigr] = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot [g]_{W^{1,2}(\Rd)}^2,
+\Ebb\Bigl[\abs{D_{k_q}f(x_h)} \cdot \abs{D_{\ell_q}f(x_j)}\Bigr] = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot [f]_{H^1(\Xset)}^2,
 \end{equation*}
@@
 \begin{equation*}
-\Ebb\Bigl[\bigl|D_{k}g(x_i)\bigr| \cdot \bigl|D_{\ell}g(x_i)\bigr|\Bigr] = O(r^{(\abs{k \cup \ell \cup i} - (2q + 1))d + 2}) \cdot [g]_{W^{1,2}(\Rd)}^2.
+\Ebb\Bigl[\bigl|D_{k}f(x_i)\bigr| \cdot \bigl|D_{\ell}f(x_i)\bigr|\Bigr] = O(r^{(\abs{k \cup \ell \cup i} - (2q + 1))d + 2}) \cdot [f]_{H^1(\Rd)}^2.
 \end{equation*}
@@
 \begin{align*}
-\Ebb \left[\bigl(D_{i}g(x_j)\bigr)^2 \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\right] &= \Ebb \left[\bigl(D_{i}g(x_j)\bigr)^2 \cdot \Pbb\bigl[\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \} |x_i,x_j\bigr]\right] \\
-& = O(r^{(\abs{k \cup \ell \cup i} - 2)d}) \cdot \Ebb\left[\bigl(D_{i}g(x_j)\bigr)^2\right] \\
-& = O(r^{(\abs{k \cup \ell \cup i} - 3)d}) \cdot \Ebb\left[\bigl(d_{i}g(x_j)\bigr)^2K_r(x_i,x_j)\right] \\
-& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot [g]_{W^{1,2}(\Xset)}^2
+\Ebb \left[\bigl(D_{i}f(x_j)\bigr)^2 \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\right] &= \Ebb \left[\bigl(D_{i}f(x_j)\bigr)^2 \cdot \Pbb\bigl[\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \} |x_i,x_j\bigr]\right] \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 2)d}) \cdot \Ebb\left[\bigl(D_{i}f(x_j)\bigr)^2\right] \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 3)d}) \cdot \Ebb\left[\bigl(d_{i}f(x_j)\bigr)^2K_r(x_i,x_j)\right] \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot [f]_{H^1(\Xset)}^2
 \end{align*}
@@
 \begin{align*}
-\Ebb \Bigl[ & \abs{D_{i}g(x_j)} \cdot \abs{D_ig(x_h)} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\Bigr] = \nonumber \\
-& \Ebb\Bigl[\abs{D_{i}g(x_j)} \cdot \abs{D_ig(x_h)} \cdot \Pbb\bigl[\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \} |x_i,x_j,x_h\bigr]\Bigr] \nonumber \\
-& = O(r^{(\abs{k \cup \ell \cup i} - 3)d}) \cdot \Ebb\Bigl[\abs{D_{i}g(x_j)} \cdot \abs{D_ig(x_h)} \Bigr] \nonumber \\
-& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2})\cdot[g]_{W^{1,2}(\Rd)}^2
+\Ebb \Bigl[ & \abs{D_{i}f(x_j)} \cdot \abs{D_if(x_h)} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\Bigr] = \nonumber \\
+& \Ebb\Bigl[\abs{D_{i}f(x_j)} \cdot \abs{D_if(x_h)} \cdot \Pbb\bigl[\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \} |x_i,x_j,x_h\bigr]\Bigr] \nonumber \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 3)d}) \cdot \Ebb\Bigl[\abs{D_{i}f(x_j)} \cdot \abs{D_if(x_h)} \Bigr] \nonumber \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2})\cdot[f]_{H^1(\Xset)}^2
 \end{align*}
@@
 \begin{align*}
-\Ebb\Bigl[ &\abs{D_{k_q}g(x_i)}\cdot{\abs{D_{\ell_q}g(x_j)}} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\Bigr] \\
-& = \Ebb\Bigl[\abs{D_{k_q}g(x_i)}\cdot{\abs{D_{\ell_q}g(x_j)}} \cdot\Pbb\bigl[G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected}|x_i,x_j,x_{k_q},x_{\ell_q}\bigr]\Bigr] \\
-& = O(r^{(\abs{k \cup \ell \cup i} - 4)d}) \cdot \Ebb\Bigl[\abs{D_{k_q}g(x_i)}\cdot{\abs{D_{\ell_q}g(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr] \\
-& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot [g]_{W^{1,2}(\Rd)}^2
+\Ebb\Bigl[ &\abs{D_{k_q}f(x_i)}\cdot{\abs{D_{\ell_q}f(x_j)}} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\Bigr] \\
+& = \Ebb\Bigl[\abs{D_{k_q}f(x_i)}\cdot{\abs{D_{\ell_q}f(x_j)}} \cdot\Pbb\bigl[G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected}|x_i,x_j,x_{k_q},x_{\ell_q}\bigr]\Bigr] \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 4)d}) \cdot \Ebb\Bigl[\abs{D_{k_q}f(x_i)}\cdot{\abs{D_{\ell_q}f(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr] \\
+& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot [f]_{H^1(\Xset)}^2
 \end{align*}
@@
 
-By Lemma~\ref{lem:leading_term_sobolev} there exists some $f_s \in \Leb^2(\Xset)$ which satisfies $\norm{f_s}_{\Leb^2(\Xset)} \leq c \norm{g}_{W^{s,2}(\Xset)}$ such that
-\begin{equation*}
-\Ebb\Bigl[\bigl(D_kg(x)\bigr)^2\Bigr] = r^{2s} \cdot \norm{f_s}_{\Leb^2(\Xset)}.
-\end{equation*} Therefore, by the law of iterated expectation 
-\begin{align*}
-\Ebb\bigl[D_kg(x_i)D_kg(x_j)\bigr] = \Ebb\Bigl[\bigl(\Ebb[D_kg(x_i)|x_i]\bigr)^2\Bigr] = r^{2s} \cdot \Ebb\Bigl[\bigl(f_s(x_i)\bigr)^2\Bigr] \leq r^{2s} \cdot c \norm{g}_{W^{s,2}(\Xset)},
-\end{align*}
-proving the claimed result.
+The result follows from the law of iterated expectation, the condition that the density be upper bounded, and equations~\eqref{eqn:leading_term_sobolev_compact_1} and \eqref{eqn:leading_term_sobolev_compact_2} in Lemma~\ref{lem:leading_term_sobolev_compact}
+\begin{equation*}
+\Ebb\bigl[D_kf(x_i)D_kf(x_j)\bigr] = \Ebb\Bigl[\bigl(\Ebb[D_kf(x_i)|x_i]\bigr)^2\Bigr] \leq p_{\max} \norm{\Ebb\bigl[D_kf\bigr]}_{\Leb^2(\Xset)}^2 \leq c r^{2s} \norm{f}_{H^s(\Xset)}^2.
+\end{equation*}
 
@@
 
-By the law of iterated expectation, we have
-\begin{align}
-\Ebb\biggl[\Bigl(d_iD_kg_m(x_j)\Bigr) \Bigl(d_iD_{\ell}g_m(x_j)\Bigr) K_r(x_i,x_j)\biggr] & = \Ebb\biggl[\Bigl(d_i\bigl(\Ebb[D_kf|x_i,x_j]\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \nonumber \\
-& = \Ebb\biggl[\Bigl(d_i\bigl(r^{s - 1}\cdot f_{s - 1} + r^sf_s(x_j)\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr]. \label{eqn:expected_difference_operators_sobolev_pf1}
-\end{align}
-where the latter equality follows from Lemma~\ref{lem:leading_term_sobolev} and $f_{s -1}$ and $f_s$ satisfy the conclusions of that Lemma, namely that $f_{s - 1} \in C^1(\Xset) \cap W^{1,2}(\Xset)$ and $f_s \in C^{0}(\Xset) \cap \Leb^2(\Xset)$, and
-\begin{equation*}
-\norm{f_{s - 1}}_{W^{1,2}(\Xset)}, \norm{f_s}_{\Leb^2(\Xset)} \leq c \norm{g}_{W^{s,2}(\Xset)}.
-\end{equation*}
-Applying these estimates inside \eqref{eqn:expected_difference_operators_sobolev_pf1}, we obtain
-\begin{align*}
-\Ebb\biggl[\Bigl(d_i\bigl(r^{s - 1}\cdot f_{s - 1} + r^s f_s(x_j)\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] & = \Ebb\biggl[\Bigl(r^{s - 1} \cdot d_if_{s - 1}(x_j) + r^s d_if_s(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \\
-& \leq 2r^{2(s - 1)} \Ebb\Bigl[\bigl(d_if_{s-1}(x_j)\bigr)^2K_r(x_i,x_j)\Bigr] + 2r^{2s}\Ebb\Bigl[\bigl(d_if_s(x_j)\bigr)^2\Bigr] \\
-& \leq 2r^{2(s - 1)} \Ebb\Bigl[\bigl(d_if_{s-1}(x_j)\bigr)^2K_r(x_i,x_j)\Bigr] + 4r^{2s} \norm{g}_{W^{s,2}(\Rd)}^2 \\
-& \leq r^{2s} \cdot c\norm{g}_{W^{s,2}(\Rd)}^2
-\end{align*}
-where the last inequality follows from Lemma~\ref{lem:expected_first_order_seminorm}.
-This concludes the proof of~\eqref{eqn:expected_difference_operators_1} and~\eqref{eqn:expected_difference_operators_2}.
+By Lemma~\ref{lem:leading_term_sobolev_compact}, we have that for $k \in (n)^{(s - 1)/2}$, there exists a function $f_{s - 1,q} \in H_0^1(\Xset)$ such that
+\begin{equation*}
+\Bigl\|\Ebb\bigl[D_kf\bigr] - r^{2{s - 1}}f_{s - 1,q}\Bigr\|_{\Leb^2(\Xset)} \leq c r^{2s} \norm{f}_{H^s(\Rd)}^2
+\end{equation*}
+and $\norm{f_{s - 1,q}}_{H^1(\Xset)} \leq c \norm{f}_{H^s(\Xset)}$. By the law of iterated expectation, and writing $\varDelta := \Ebb\bigl[D_kf\bigr] - r^{2{s - 1}}f_{s - 1,q}$, we have
+\begin{align*}
+\Ebb\biggl[\Bigl(d_iD_kf(x_j)\Bigr) \Bigl(d_iD_{\ell}f(x_j)\Bigr) K_r(x_i,x_j)\biggr] & =  \Ebb\biggl[\Bigl(d_i\bigl(\Ebb[D_kf|x_i,x_j]\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \nonumber \\
+& = \Ebb\biggl[\Bigl(d_i\bigl(r^{s - 1}\cdot f_{s - 1,q} + \varDelta \bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \\
+& \leq 2r^{2(s - 1)}\Ebb\biggl[\Bigl(d_i\bigl(f_{s - 1,q}\bigr)\Bigr)^2K_r(x_i,x_j)\biggr]  + 2 \Ebb \biggl[\Bigl(d_i\varDelta(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \\
+& \leq 2r^{2(s - 1)}\Ebb\biggl[\Bigl(d_i\bigl(f_{s - 1,q}\bigr)\Bigr)^2K_r(x_i,x_j)\biggr] + 2p_{\max}^2\norm{\varDelta}_{\Leb^2(\Xset)}^2 \\
+& \leq c r^{2s} \norm{f}_{H^s(\Xset)}^2
+\label{eqn:expected_difference_operators_sobolev_pf1}
+\end{align*}
+where the last inequality follows by Lemma~\ref{lem:expected_first_order_seminorm}.
 
@@
 Select a bounded open set $V$ such that $U \subset \subset V$ ($U$ is compactly contained in $V$).
-Then there exists a bounded linear operator $E: W^{1,2}(U) \to W^{1,2}(\Rd)$ such that for each $u \in W^{1,2}(U)$:
+Then there exists a bounded linear operator $E: H^1(U) \to H^1(\Rd)$ such that for each $u \in H^1(U)$:
 	\begin{enumerate}
@@
 		\begin{equation*}
-		\norm{Eu}_{W^{1,2}(\Rd)} \leq C \norm{u}_{W^{1,2}(\Rd)}
+		\norm{Eu}_{H^1(\Rd)} \leq C \norm{u}_{H^1(\Rd)}
 		\end{equation*}
@@
 For Lemmas~\ref{lem:expected_first_order_seminorm} - \ref{lem:expected_first_order_seminorm_3}, we will assume that $K$ is a kernel function compactly supported on $B(0,1)$ and upper bound $K(x) \leq K_{\max}$.
+Additionally, we note that although each Lemma assumes $g \in C^1(\Xset)$, in each case if we assume merely $g \in H^1(\Xset)$ the results still hold due to the density of $C^{\infty}$ in $H^1(\Xset)$.
 
@@
 	\begin{equation*}
-	\Ebb\Bigl[(g(x_j) - g(x_i))^2K_r(x_i,x_j)\Bigr] \leq c K_{\max} p_{\max}^2 r^2 [g]_{W^{1,2}(\Xset)}^2
+	\Ebb\Bigl[(g(x_j) - g(x_i))^2K_r(x_i,x_j)\Bigr] \leq c K_{\max} p_{\max}^2 r^2 [g]_{H^1(\Xset)}^2
 	\end{equation*}
@@
 	\begin{equation*}
-	[h]_{W^{1,2}(\Rd)} \leq c[g]_{W^{1,2}(\Xset)}
+	[h]_{H^1(\Rd)} \leq c[g]_{H^1(\Xset)}
 	\end{equation*}
@@
 	\begin{equation*}
-	\Ebb\Bigl[\abs{D_if(x_h)}\cdot\abs{D_if(x_j)} \Bigr] \leq c K_{\max}^3 p_{\max}^3 r^2 [f]_{W^{1,2}(\Xset)}^2
+	\Ebb\Bigl[\abs{D_if(x_h)}\cdot\abs{D_if(x_j)} \Bigr] \leq c K_{\max}^3 p_{\max}^3 r^2 [f]_{H^1(\Xset)}^2
 	\end{equation*}
@@
 	\begin{equation*}
-	[g]_{W^{1,2}(\Rd)} \leq c[f]_{W^{1,2}(\Xset)}
+	[g]_{H^1(\Rd)} \leq c[f]_{H^1(\Xset)}
 	\end{equation*}
@@
 	\begin{equation*}
-	\Ebb\Bigl[\abs{D_{k}f(x_i)}\cdot{\abs{D_{\ell}f(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr] \leq c K_{\max} p_{\max}^2 r^{2 + d} [f]_{W^{1,2}(\Rd)}^2
+	\Ebb\Bigl[\abs{D_{k}f(x_i)}\cdot{\abs{D_{\ell}f(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr] \leq c K_{\max} p_{\max}^2 r^{2 + d} [f]_{H^1(\Rd)}^2
 	\end{equation*}
@@
 
+Lemma~\ref{lem:leading_term_sobolev_compact} supplies an equivalent result when $f \in H_0^{s}(\Xset)$.
+In this Lemma, we write $\Ebb[D_kf]:\Xset \to \Reals$ for the mapping $x \mapsto \Ebb[D_kf(x)]$.
+We will also denote $U_r = \Xset \setminus \Xset_{r}$.
 \begin{lemma}
-	\label{lem:leading_term_sobolev}
-	Let $k \in [n]^q$ for some $q \geq 1$, let $K_r$ be a second order kernel, and suppose $f \in C^{s}(\Xset) \cap W_{\sigma}^{s,2}(\Xset)$ for some $s \in \mathbb{N}$ and $\sigma > 0$.
-Let $p$ be a density satisfying $p \in C^{0}(\Xset;p_{\max})$ if $s = 0$, and otherwise $p \in C^{s - 1}(\Xset;p_{\max})$ for some $p_{\max} > 0$.
-For any $qr < \sigma$, there exist functions $f_{\ell}$ for $\ell = 2q,\ldots,s - 1$ and $f_s$ satisfying $f_{\ell} \in C^{s - \ell}(\Xset) \cap W_{\sigma - qr}^{s - \ell,2}(\Xset)$ and
-	\begin{equation*}
-	\norm{f_{\ell}}_{W^{s - \ell,2}(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}
-	\end{equation*}
-	for some constant $c$ which depends only on $\sigma$, $d$, $\Xset$ and $p_{\max}$ such that
-	\begin{equation}
-	\label{eqn:leading_term_sobolev}
-	\Ebb(D_kf(x)) =
-	\begin{cases*}
-	\sum_{\ell = 2q}^{s} f_{\ell}(x) r^{\ell} ,~~& \textrm{if $2q < s$} \\
-	r^s \cdot f_{s}(x),~~& \textrm{if $2q \geq s$}
-	\end{cases*}
-	\end{equation}
-	for any $x \in \Xset$.
+	\label{lem:leading_term_sobolev_compact}
+	Fix integers $s \geq 0$ and $q \geq 1$, and an index vector $k \in (n)^q$.
+Suppose that $p \in C^0(\Xset;p_{\max})$, and additionally that $p \in C^{s-1}(\Xset;p_{\max})$ if $s \geq 2$.
+Then there exists an $r' > 0$ such that for all $0 < r < r'$, the following statements hold for all $f \in H_0^{s}(\Xset)$:
+	\begin{itemize}
+		\item The expected difference operator $\Ebb[D_kf]$ belongs to $\Leb^2(U_{qr})$, with norm
+		\begin{equation}
+		\label{eqn:leading_term_sobolev_compact_1}
+		\Bigl\|\Ebb\bigl[D_kf\bigr]\Bigr\|_{\Leb^2(U_{qr})} \leq c r^s \norm{f}_{H^s(\Xset)}
+		\end{equation}
+		\item If $2q \geq s$, then additionally $\Ebb[D_kf]$ belongs to $\Leb^2(\Xset_{qr})$, with norm
+		\begin{equation}
+		\label{eqn:leading_term_sobolev_compact_2}
+		\Bigl\|\Ebb\bigl[D_kf\bigr]\Bigr\|_{\Leb^2(\Xset_{qr})} \leq c r^s \norm{f}_{H^s(\Xset)}.
+		\end{equation}
+		Otherwise there exist functions $f_{\ell} \in H_0^{s - \ell}(\Xset)$, which additionally satisfy
+		\begin{equation}
+		\label{eqn:leading_term_sobolev_compact_3}
+		\norm{f_{\ell}}_{H^{s - \ell}(\Xset)} \leq c \norm{f}_{H^s(\Xset)}
+		\end{equation}
+		for each $\ell = 2q,\ldots,s-1$, such that
+		\begin{equation}
+		\label{eqn:leading_term_sobolev_compact_4}
+		\Bigl\|\Ebb\bigl[D_kf\bigr] - \sum_{\ell = 2q}^{s - 1} r^{\ell} f_{\ell}\Bigr\|_{\Leb^2(\Xset_{qr})} \leq c r^s \norm{f}_{H^s(\Xset)}
+		\end{equation}
+	\end{itemize}
 \end{lemma}
 \begin{proof}
-	We proceed by induction on $q$.
+	One can interpret the conclusions of Lemma~\ref{lem:leading_term_sobolev_compact} as demonstrating that expected difference operators behave similarly to derivatives over $H_0^{s}(\Xset)$.
+The proof of Lemma~\ref{lem:leading_term_sobolev_compact} is therefore naturally centered on taking Taylor expansions, but in order to do this, we must relate $f$ to a function $g$ which has classical derivatives.
+
+	Since $f \in H_0^s(\Xset)$, there exists a sequence $(f_m) \subset C_c^s(\Xset)$ such that $\norm{f_m - f}_{H^s(\Xset)} \to 0$ as $m \to \infty$ (Indeed $f_m$ will be smooth for each $m$, but we will not need that fact.) Picking $m$ large enough so that
+	\begin{equation*}
+	\norm{f_m - f}_{H^s(\Xset)} \leq r^s \norm{f}_{H^s(\Xset)}
+	\end{equation*}
+	we have that for any $\wt{f} \in \Leb^2(\Xset)$,
+	\begin{align*}
+	\norm{\Ebb\Bigl[D_kf\Bigr] - \wt{f}}_{\Leb^2(\Xset)} & \leq \norm{\Ebb\Bigl[D_kf_m\Bigr] - \wt{f}}_{\Leb^2(\Xset)} + \norm{\Ebb\Bigl[D_k(f - f_m)\Bigr]}_{\Leb^2(\Xset)} \\
+	& \leq \norm{\Ebb\Bigl[D_kf_m\Bigr] - \wt{f}}_{\Leb^2(\Xset)} + c \norm{f - f_m}_{\Leb^2(\Xset)} \\
+	& \leq \norm{\Ebb\Bigl[D_kf_m\Bigr] - \wt{f}}_{\Leb^2(\Xset)} + c r^s \norm{f}_{H^s(\Xset)}
+	\end{align*}
+	Since $f_m \in C_c^s(\Xset)$, it can be continuously extended to $g: \Rd \to \Reals,~ g \in C_c^s(\Xset)$ by taking $g(x) = 0$ for all $x \in \Rd \setminus \Xset$, such that $\norm{g}_{H^s(\Rd)} = \norm{f_m}_{H^s(\Xset)}$ and $g = f$ everywhere on $\Xset$.
+Therefore $\Ebb[D_kg] =\Ebb[D_kf_m]$, and it suffices to prove the estimates~\eqref{eqn:leading_term_sobolev_compact_1}-\eqref{eqn:leading_term_sobolev_compact_4} hold with respect to $g$.
+
+	Before we do so, let us establish some notation. When $s \geq 1$, since $g \in C_c^s(\Rd)$ it admits a Taylor expansion of the form
+	\begin{equation*}
+	g(y) =  \sum_{\abs{\alpha} = 0}^{s - 1} g^{(\alpha)}(x) (z - x)^{\alpha} + \sum_{\abs{\alpha} = s} (y - x)^{\alpha} G_{\alpha}(x,y),
+	\end{equation*}
+	for any $y,x \in \Rd$. When $s \geq 2$, since $p \in C_c^{s - 1}(\Xset)$ it also admits a Taylor expansion,
+	\begin{equation*}
+	p(y) = \sum_{\abs{\beta} = 0}^{s - 2} p^{\beta}(x) (y - x)^{\beta} + \sum_{\abs{\beta} = s - 1} (x - y)^s P_{\beta}(x,y).
+	\end{equation*}
+	for any $y,x \in \Xset$.
+	In both cases we use the integral form of the remainders:
+	\begin{align*}
+	G_{\alpha}(x,y) & = \int_{0}^{1} g^{(\alpha)}\bigl(x + t(y - x)\bigr)(1 - t)^{\abs{\alpha}} \,dt \\
+	P_{\beta}(x,y) & = \int_{0}^{1} p^{(\beta)}\bigl(x + t(y - x)\big) (1 - t)^{\abs{\beta}}  \,dt 
+	\end{align*}
+	where by Rademacher's Theorem $g^{(\alpha)}$ and $p^{(\beta)}$ exist almost everywhere, and the preceding integrals are therefore well defined.
+
+	It will also be helpful to introduce some notation. For $G: \Xset \times \Xset \to \Reals$, let
+	\begin{align*}
+	\Bigl(\Ebb_{\alpha}[G]\Bigr)(x) & := \int_{\Xset} (y - x)^{\alpha} G(x,y) K_r(y,x) p(y) \,dy,~~ && \Ebb_{\alpha}(x) := \Bigl(\Ebb_{\alpha}[1]\Bigr)(x) \\
+	\Bigl(\Ibb_{\alpha}[G]\Bigr)(x) & := \int_{\Xset} (y - x)^{\alpha} G(x,y) K_r(y,x) \,dy,~~ && \Ibb_{\alpha}(x)  := \Bigl(\Ibb_{\alpha}[1]\Bigr)(x)
+	\end{align*}
+	Additionally we define
+	\begin{equation*}
+	I_{\alpha} := \int z^{\alpha} K\bigl(\norm{z}\bigr) \,dz.
+	\end{equation*}
+	and note that when $B(x,r) \subset \Xset$, the following two facts are true: first, that $\Ibb_{\alpha,x} = r^{\abs{\alpha}} I_{\alpha}$, and second that $I_{\alpha} = 0$ when $\abs{\alpha} = 1$.
+
+	We begin with $s = 0$. When $q = 1$, we have
+	\begin{align*}
+	\norm{\Ebb\bigl[D_kg\bigr]}_{\Leb^2(\Xset)} & = \int_{\Xset} \biggl[\int_{\Xset} \Bigl(g(y) - g(x)\Bigr)K_r(y,x) p(y) \,dy \biggr]^2 \,dx \\
+	& \leq p_{\max}^2 \int_{\Rd} \biggl[\int_{\Rd} \Bigl(\abs{g(y)} + \abs{g(x)}\Bigr)K_r(y,x) \,dy \biggr]^2 \,dx \\
+	& \leq K_{\max}^2 p_{\max}^2 \int_{\Rd} \biggl[\int_{B(0,1)} \abs{g(zr + x)} + \abs{g(x)} \,dz \biggr]^2 \,dx
+	\end{align*}
+	and the statement follows by Lemma~\ref{lem:remainder_term}. The same result holds (up to different constants) for $s = 0$ and general $q$ by induction.
+
+	For $s \geq 1$, we first show the desired estimate over $U_{qr}$.
+
+	\subsubsection{Boundary region}
+
+	Take $q = 1$, and $k \in [n]$. We begin by relating the $\Leb^2$ norm of $\Ebb[D_kg]$ over $U_r$ to the $\Leb^2$ norm of $g$ over $U_{2r}$, as follows:
+	\begin{align*}
+	\Bigl\|\Ebb\bigl[D_kg\bigr]\Bigr\|_{\Leb^2(U_r)}^2 & = \int_{U_r} \biggl[\int_{\Xset} \bigl(g(y) - g(x)\bigr)K_r(x,y) p(y) \,dy \biggr]^2 \,dx \\
+	& \leq p_{\max}^2 \int_{U_r} \biggl[\int_{\Xset} \bigl(\abs{g(y)} +  \abs{g(x)}\bigr)K_r(x,y) \,dy \biggr]^2 \,dx \\
+	& \overset{(i)}{\leq}  p_{\max}^2 K_{\max}^2 \int_{U_r} \biggl[\int_{B(0,1) \cap (\Xset - x)/r} \bigl(\abs{g(zr + x)} +  \abs{g(x)}\bigr)\,dz \biggr]^2 \,dx \\
+	& \overset{(ii)}{\leq} 2 p_{\max}^2 K_{\max}^2 \int_{U_r} \nu_d^2 \bigl(g(x)\bigr)^2 \,dx + 2 p_{\max}^2 K_{\max}^2 \nu_d \int_{U_r}\biggl[\int_{B(0,1) \cap (\Xset - x)/r} \bigl(g(zr + x)\bigr)^2\,dz\biggr] \,dx \\
+	& \leq 2 p_{\max}^2 K_{\max}^2 \int_{U_r} \nu_d^2 \bigl(g(x)\bigr)^2 \,dx + 2 p_{\max}^2 K_{\max}^2 \nu_d \int_{B(0,1)} \int_{U_r} \bigl(g(zr + x)\bigr)^2\,dz \,dx \\
+	& \leq 4 p_{\max}^2 K_{\max}^2 \nu_d^2 \norm{g}_{\Leb^2(U_{2r})}^2
+	\end{align*}
+	where $(i)$ follows from change of variables and $(ii)$ from Young's and Jensen's inequality. Reasoning by induction, we see that it suffices to show that
+	\begin{equation*}
+	\norm{g}_{\Leb^2(U_{(q + 1)r})}^2 \leq c r^{2s} \norm{g}_{H^s(\Xset)}^2
+	\end{equation*}
+	to establish~\eqref{eqn:leading_term_sobolev_compact_1}.
+The previous inequality is established in Lemma~\ref{lem:boundary_term_sobolev} for all $r > 0$ sufficiently small, and we have therefore proved the desired estimate over the boundary.
+
+	\subsubsection{Interior region}
+
+	To show the desired bounds on $\Xset_{qr}$ when $s \geq 1$, we reason by induction on $q$.
 
 	\paragraph{Base case.}
-	We begin with the base case of $q = 1$.
-When $s = 0$, we need to show that $\Ebb[D_kf(x)] = f_s(x)$ for some $f_s \in C^{0}(\Xset)$ satisfying $\mathrm{supp}(f_s) \subset X_{\sigma - r}$ and $\norm{f_s}_{\Leb^2(\Xset)} \leq c\norm{f}_{W^{s,2}(\Xset)}$.
-But
-	\begin{align*}
-	\Ebb\Bigl[D_kf(x)\Bigr] & = \int f(z) K_r(x,z) p(z) \,dz  - f(x)\Ebb\bigl[K_r(x_k,x)\bigr] \\
-	& = \int f(yr + x) K\bigl(\norm{y}\bigr) p(yr + z) \,dy - f(x)\Ebb\bigl[K_r(x_k,x)\bigr],
-	\end{align*}
-	whence the boundedness $f_s \in C^{0}(\Xset)$ follows from the boundedness of $f, K$ and $p$.
-We now analyze each term in the above difference.
-First, we have that $f(x)\Ebb[K_r(x_k,x)] = 0$ unless $x \in \Xset_{\sigma}$, and
-	\begin{equation*}
-	\norm{f\Ebb[K_r(x_k,\cdot)]}_{\Leb^2(\Xset)} \leq p_{\max} \norm{f}_{\Leb^2(\Xset)}.
-	\end{equation*}
-	Moreover, since $K$ is compactly supported on $B(0,1)$, the integral $\int f(yr + x) K(\norm{y}) p(yr + x) \,dy = 0$ unless $x \in \Xset_{\sigma - r}$. Since
-	\begin{equation*}
-	\biggl\|\int f(yr + x) K(\norm{y}) p(yr + x) \,dy\biggr\|_{\Leb^2(\Xset)} \leq p_{\max} K_{\max} \norm{f}_{\Leb^2(\Xset)},
-	\end{equation*}
-	the claim follows.
-
-	Now, when $s \geq 1$ since $f \in C^{s}(\Xset)$ it admits a Taylor expansion of the following form for all $x,z \in \Xset$:
-	\begin{equation*}
-	f(z) = \sum_{\abs{\alpha} < s} \frac{f^{(\alpha)}(x)}{\alpha!} (x - z)^{\alpha} + \frac{1}{(s - 1)!}\sum_{\abs{\alpha} = s} (x - z)^{\alpha} \int_{0}^{1}(1 - t)^{s - 1} f^{(\alpha)}(x + t(z - x)) \,dt
-	\end{equation*}
-	$f^{(\alpha)} \in C^{s - \abs{\alpha}}(\Xset)$ additionally satisfies
-	\begin{equation*}
-	\norm{f^{(\alpha)}}_{W^{s - \abs{\alpha},2}(\Xset)} \leq \norm{f}_{W^{s,2}(\Xset)} ~~\textrm{and}~~ \mathrm{supp}\bigl(f^{(\alpha)}\bigr) \subset {\Xset}_{\sigma}.
-	\end{equation*}
-	Replacing $f$ by its Taylor expansion inside the expected first order difference operator $\Ebb(D_kf(x))$ we have
-	\begin{equation}
-	\label{eqn:leading_term_sobolev_pf1}
-	\Ebb(D_kf(x)) = \sum_{1 \leq \abs{\alpha} < s} \frac{f^{(\alpha)}(x)}{\alpha!} E_{\alpha,x} + \frac{1}{(s - 1)!}\sum_{\abs{\alpha} = s} \int_{0}^{1}(1 - t)^{s - 1} E_{\alpha,x,t}(f)  \,dt 
-	\end{equation}
-	where we use the notation $E_{\alpha,x} := \Ebb\left[(x - x_k)^{\alpha}K_r(x_k,x)\right]$ and  $E_{\alpha,x,t}(f) := \Ebb\bigl[f^{(\alpha)}(x + t(x_k - x)) (x_k - x)^{\alpha} K_r(x_k,x)\bigr]$.
-
-	By a change of variables, we have
-	\begin{align*}
-	E_{\alpha,x,t}(f) & = \frac{1}{r^d}\int_{\Rd} f^{(\alpha)}\bigl(x + t(z - x)\bigr) (z - x)^{\alpha}  K\biggl(\frac{\norm{z - x}}{r}\biggr) p(z) \,dz \\
-	& = r^s \int_{\Rd} y^{\alpha} f^{(\alpha)}\bigl(x + tyr) K\bigl(\norm{y}\bigr) p(yr + x) \,dy
-	\end{align*}
-	and as a result the remainder term in \eqref{eqn:leading_term_sobolev_pf1} reduces to
-	\begin{align*}
-	\frac{1}{(s - 1)!}\sum_{\abs{\alpha} = s} \int_{0}^{1}(1 - t)^{s - 1} E_{\alpha,x,t}(f)  \,dt & = \frac{r^s}{(s - 1)!}\sum_{\abs{\alpha} = s} \int_{0}^{1} \int_{\Rd} (1 - t)^{s - 1}  y^{\alpha} f^{(\alpha)}\bigl(x + tyr) K\bigl(\norm{y}\bigr) p(yr + x) \,dy \,dt \\
-	& =: r^s g_s(x) ~~\textrm{for $g_s \in C^{0}(\Xset) \cap \Leb_{\sigma - r}^2(\Xset)$},
-	\end{align*}
-	where additionally $\norm{g_s}_{\Leb^2(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}$ by Lemma~\ref{lem:remainder_term}.
-When $s = 1$, there are no multiindices $1 < \abs{\alpha} < s$, and so only the remainder term in \eqref{eqn:leading_term_sobolev_pf1} is non-zero; thus we have shown~\eqref{eqn:leading_term_sobolev} when $q = 1$ and $s = 1$.
-
-	When $s \geq 2$, we can analyze $E_{\alpha,x}$ using a Taylor expansion of $p$.  For any $x,z \in \Xset$, we have
-	\begin{equation*}
-	p(z) = \sum_{\abs{\beta} < s  - 1} \frac{p^{(\beta)}(x)}{\beta!} (x - z)^{\beta} + \frac{1}{(s - 2)!}\sum_{\abs{\beta} = s - 1} (x - z)^{\beta} \int_{0}^{1}(1 - t)^{s - 2} p^{(\beta)}(x + t(z - x)) \,dt
-	\end{equation*}
-	and $p^{(\beta)} \in C^{s - \abs{\beta} - 1}(\Xset)$ additionally satisfies
-	\begin{equation*}
-	\norm{p^{(\beta)}}_{C^{s - \abs{\beta} - 1}(\Xset)} \leq p_{\max}.
-	\end{equation*}
-	Replacing $p$ by its Taylor expansion, we analyze the term $E_{\alpha,x}$. Note that by assumption $\sigma > r$, and therefore for any $x \in \Xset_{\sigma}$,
-	\begin{equation*}
-	\int_{\Xset} (x - z)^{\alpha + \beta} K_r(x,z) \,dz = \int_{\Rd} (x - z)^{\alpha + \beta} K_r(x,z) \,dz = r^{\abs{\alpha} + \abs{\beta}} \int_{\Rd} y^{\alpha + \beta} K\bigl(\norm{y}\bigr) \,dz
-	\end{equation*} 
-	and similarly
-	\begin{equation*}
-	\int_{\Xset} \int_{0}^{1} (1 - t)^{s - 2} (x - z)^{\alpha + \beta}  p^{(\beta)}(x + t(z - x)) K_r(x,z) \,dz \,dt = r^{\abs{\alpha} + \abs{\beta}}\int_{\Rd} \int_{0}^{1} (1 - t)^{s - 2} y^{\alpha + \beta}  p^{(\beta)}(x + try) K\bigl(\norm{y}\bigr) \,dy \,dt
-	\end{equation*}
-	Therefore for any such $x \in \Xset_{\sigma}$,
-	\begin{align*}
-	E_{\alpha,x} & = \int_{\Xset} (x - z)^{\alpha} K_r(x,z) p(z) \,dz \\
-	& =  \sum_{\abs{\beta} = 0}^{s - 2} \frac{p^{(\beta)}(x)}{\beta!} \int_{\Xset} (x - z)^{\alpha + \beta}K_r(x,z)\,dz ~~+ \\
-	& ~~~\frac{1}{(s - 2)!}\sum_{\abs{\beta} = s - 1} \int_{\Xset} \int_{0}^{1} (1 - t)^{s - 2} (x - z)^{\alpha + \beta}  p^{(\beta)}(x + t(z - x)) K_r(x,z) \,dz \,dt \\
-	& = \sum_{\abs{\beta} = 0}^{s - 2} r^{\abs{\alpha} + \abs{\beta}} \frac{p^{(\beta)}(x)}{\beta!} \int_{\Rd} y^{\alpha + \beta}K(y)\,dy +~ \frac{r^{\abs{\alpha} + s - 1}}{(s - 2)!} \sum_{\abs{\beta} = s - 1} \int_{\Rd} \int_{0}^{1} (1 - t)^{s - 2} y^{\alpha + \beta} p^{(\beta)}(x + ytr) K\bigl(\norm{y}\bigr) \,dy \,dt \\
-	& =: \sum_{\abs{\beta} = 0}^{s - 2} \1\{\abs{\alpha} + \abs{\beta} > 1\}r^{\abs{\alpha} + \abs{\beta}} p_{\abs{\beta}}(x) + r^s p_{s - 1}(x) ~~\textrm{for $p_{\abs{\beta}} \in C^{\infty}(\Xset), \norm{p_{\abs{\beta}}}_{C^{s - \abs{\beta} - 1}(\Xset)} \leq c \norm{p}_{C^{s - 1}(\Xset)}$},
-	\end{align*}
-	where the last line follows since $K$ is a 2nd order kernel. On the other hand when $x \not\in \Xset_{\sigma}$, the derivatives $f^{(\alpha)}(x) = 0$ for each $\alpha = 1,2,\ldots,s-1$. 
-	As a result, by substituting our expressions for $E_{\alpha,x}$ and $E_{\alpha,x,t}(f)$ back into~\eqref{eqn:leading_term_sobolev_pf1}, we obtain that for all $x \in \Xset$,
-	\begin{align*}
-	\Ebb\bigl[D_kf(x)\big] & = \sum_{\abs{\alpha} = 1}^{s - 1} \sum_{\abs{\beta} = 0}^{s-2} \1\{\abs{\alpha} + \abs{\beta} > 1\} r^{\abs{\alpha} + \abs{\beta}} \frac{f^{(\alpha)}(x)p_{\abs{\beta}}(x)}{\alpha!} + r^s \biggl(\sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)p_{s - 1}(x)}{\alpha!} + g_s(x)\biggr) \\
-	& = \sum_{\ell = 2}^{s - 1} r^{\ell} \sum_{ \substack{\abs{\alpha} + \abs{\beta} = \ell, \\ \abs{\beta} > 0} }   \frac{f^{(\alpha)}(x)p_{\abs{\beta}}(x)}{\alpha!} + r^s \biggl(\sum_{\abs{\alpha} = 1}^{s - 1} \frac{f^{(\alpha)}(x)p_{s - 1}(x)}{\alpha!} + g_s(x)\biggr) \\
-	& =: \sum_{\ell = 2}^{s - 1} r^{\ell} f_{\ell}(x) + r^s f_s(x).
-	\end{align*}
-	We have already shown $g_s \in C^{0}(\Xset) \cap \Leb_{\sigma - r}^2(\Xset)$.
-Since additionally $f^{(\alpha)} \in C^{s - \abs{\alpha}}(\Xset) \cap W_{\sigma}^{s - \abs{\alpha},2}(\Xset)$ and $p_{\abs{\beta}} \in C^{s - 1 - \abs{\beta}}(\Xset)$, the products
-	\begin{equation*}
-	f^{(\alpha)} \cdot p_{\abs{\beta}} \in W_{\sigma}^{m,2}(\Xset) \cap C^{m}(\Xset) ~~\textrm{for $m = \min\{s - \abs{\alpha}, s - 1 - \abs{\beta}\}$.}
-	\end{equation*}
-	Finally, note that for $\abs{\alpha} + \abs{\beta} = \ell$ the inequality $s - \ell \leq m$ holds.
-Therefore, $f_{\ell} \in W_{\sigma}^{s - \ell,2}(\Xset) \cap C^{s - \ell}(\Xset)$ for each $\ell = 2,3,\ldots,s$, and
-	\begin{equation*}
-	\norm{f^{(\alpha)}p_{\abs{\beta}}}_{W^{s - \ell,2}(\Rd)} \leq \norm{f^{(\alpha)}p_{\abs{\beta}}}_{W^{m,2}(\Rd)} \leq p_{\max} \norm{f^{(\alpha)}}_{W^{s - \abs{\alpha},2}(\Xset)} \leq p_{\max} \norm{f}_{W^{s,2}(\Rd)}
-	\end{equation*} 
-	which finishes the proof of Lemma~\ref{lem:leading_term_sobolev} when $q = 1$.
-
-	\paragraph{Induction Step.}
-
-	Now, we assume \eqref{eqn:leading_term_sobolev} holds for all $k \in [n]^q$, and prove the desired estimate on $\Ebb[D_{k}D_jf(x)]$ for each $j \in [n]$.
-We will build a proof piece-by-piece, depending on the relative size of $s$ and $q$.
-
-	If $s \leq 2$, by hypothesis there exists $f_s \in C^{0}(\Xset) \cap \Leb_{\sigma - r}^{2}(\Xset)$ satisfying
-	\begin{equation*}
-	\norm{f_s}_{\Leb^2(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}
-	\end{equation*}
-	such that for any $z \in \Xset$,
-	\begin{equation*}
-	\Ebb\Bigl[D_jf(z)\Bigr] = r^s f_s(z).
-	\end{equation*}
-	Therefore by the law of iterated expectation,
-	\begin{equation*}
-	\Ebb\Bigl[D_kD_jf(x)\Bigr] = \Ebb\Bigl[D_k(\Ebb[D_jf])(x)\Bigr] = \Ebb\Bigl[D_k\bigl(r^s f_s\bigr)(x)\Bigr] = r^s \Ebb\Bigl[D_kf_s(x)\Bigr]
-	\end{equation*}
-	Seeing as $f_s \in C^{0}(\Xset) \cap \Leb_{\sigma - r}^2(\Xset)$, we may apply the inductive hypothesis to obtain that $\Ebb[D_kf_s(x)] \in C^0(\Xset) \cap \Leb_{\sigma - (q + 1)r}^2(\Xset)$, and additionally
-	\begin{equation}
-	\label{eqn:leading_term_sobolev_pf3}
-	\biggl\|\Ebb\Bigl[D_kf_s(x)\Bigr]\biggr\|_{\Leb^2(\Xset)} \leq c \norm{f_s}_{\Leb^2(\Xset)}.
-	\end{equation}
-	We have therefore established~\eqref{eqn:leading_term_sobolev} for all $q$ in the case when $s \leq 2$.
-
-	Otherwise, when $s \geq 3$ by hypothesis there exist functions $f_{\ell} \in C^{s - \ell}(\Xset) \cap W_{\sigma - r}^{s - \ell,2}(\Xset)$ for each $\ell = 2,\ldots,s$ satisfying
-	\begin{equation*}
-	\norm{f_{\ell}}_{W^{s - \ell,2}(\Xset)} \leq c \norm{f}_{W^{s,2}(\Xset)}
-	\end{equation*}
-	such that for any $z \in \Xset$,
-	\begin{equation*}
-	\Ebb(D_jf(z)) = \sum_{\ell = 2}^{s} r^{\ell} f_{\ell}(z).
-	\end{equation*}
-	By the law of iterated expectation
-	\begin{align*}
-	\Ebb\Bigl[D_{k}D_jf(x)\Bigr] & = \Ebb\Bigl[D_k\bigl(\Ebb[D_jf]\bigr)(x)\Bigr] \\
-	& = \Ebb\biggl[D_k\Bigl(\sum_{\ell = 2}^{s - 1} r^{\ell} f_{\ell} + r^s f_s\Bigr)(x)\biggr] \\
-	& = \sum_{\ell = 2}^{s - 1} r^{\ell} \cdot \Ebb\Bigl[D_kf_{\ell}(x)\Bigr] + r^s \Ebb\Bigl[D_kf_s(x)\Bigr].
-	\end{align*}
-	Recalling that $\Ebb\Bigl[D_kf_s(x)\Bigr]$ satisfies~\eqref{eqn:leading_term_sobolev_pf3}, we now apply the inductive hypothesis to $\Ebb(D_kf_{\ell}(x))$ for each $\ell = 2,\ldots,s-1$, to prove~\eqref{eqn:leading_term_sobolev}.
-
-	First we consider the case when $2(q + 1) \geq s$.
-Note that $2q \geq s - \ell$ for each $\ell \geq 2$.
-Therefore by hypothesis, for each $\ell = 2,\ldots,s - 1$ the expectation $\Ebb[D_kf_{\ell}(x)] = r^{s - \ell}f_{\ell,s}(x)$ for some $f_{\ell,s} \in C^{s - \ell}(\Xset) \cap \Leb_{\sigma - (q + 1)r}^2(\Rd)$ satisfying
-	\begin{equation}
-	\label{eqn:leading_term_sobolev_pf4}
-	\norm{f_{\ell,s}}_{\Leb^2(\Xset)} \leq c\norm{f_{\ell}}_{W^{s - \ell,2}(\Xset)} \leq c\norm{f}_{W^{s,2}(\Xset)}
-	\end{equation}
-	and as a result
-	\begin{equation*}
-	\sum_{\ell = 2}^{s - 1} r^{\ell} \cdot \Ebb(D_kf_{\ell}(x)) = r^s \sum_{\ell = 2}^{s - 1} f_{\ell,s}(x)
-	\end{equation*}
-	establishing that the second part of~\eqref{eqn:leading_term_sobolev} holds for all $q$ and all $s \leq 2(q + 1)$.
-
-	Otherwise $2(q + 1) < s - 1$. For each $\ell = 2,\ldots, s - 1$, if additionally  $2q \leq s - \ell - 1$, then by hypothesis
-	\begin{equation*}
-	\Ebb(D_kf_{\ell}(x)) = \sum_{m = 2q}^{s - \ell - 1} r^{m} \cdot f_{\ell, \ell + m}(x) + r^{s - {\ell}} f_{\ell,s}
-	\end{equation*}
-	where $f_{\ell,\ell + m} \in C^{s - (\ell + m)}(\Xset) \cap W_{\sigma - (q + 1)r}^{s - (\ell + m),2}(\Xset)$ satisfies
-	\begin{equation*}
-	\norm{f_{\ell,\ell + m}}_{W^{s - (\ell + m),2}(\Rd)} \leq c\norm{f_{\ell}}_{W^{s - \ell}} \leq c\norm{f}_{W^{s,2}(\Rd)}.
-	\end{equation*}
-	On the other hand if $2s > s - \ell - 1$, then
-	\begin{equation*}
-	\Ebb\bigl[D_kf_{\ell}(x)\bigr] = r^s f_{\ell,s}
-	\end{equation*}
-	for some $f_{\ell,s} \in C^0(\Xset) \cap \Leb_{\sigma - (q + 1)r}^{2}(\Xset)$ which additionally satisfies \eqref{eqn:leading_term_sobolev_pf4}. Therefore,
-	\begin{align*}
-	\sum_{\ell = 2}^{s - 1}  r^{\ell} \cdot \Ebb(D_kf_\ell(x)) & = \sum_{\ell = 2}^{s - 1 - 2q} r^{\ell} \cdot \left\{\sum_{m = 2q}^{s - \ell - 1} r^m \cdot f_{\ell,\ell + m}(x) + r^{s - \ell} \cdot f_{\ell,s}(x) \right\} + \sum_{\ell = s - 1 - 2q}^{s - 1} r^s \cdot f_{\ell,s}(x) \\
-	& = \sum_{\ell = 2}^{s - 1 - 2q} r^{\ell} \cdot \left\{\sum_{m = 2q}^{s - \ell - 1} r^m \cdot f_{\ell, \ell + m}(x)\right\}  + r^{s}\sum_{\ell = 2}^{s - 1}f_{\ell,s}(x) \\
-	& = \sum_{m = 2q}^{s - 3} \sum_{\ell = 2}^{s - m - 1} r^{m + \ell} \cdot f_{\ell, \ell + m}(x) + r^{s}\sum_{\ell = 2}^{s - 1}f_{\ell,s}(x).
-	\end{align*}
-	Rewriting the first sum in the final equation as a sum over $\ell + m = 2(q + 1),\ldots, s - 1$ establishes~\eqref{eqn:leading_term_sobolev}.
+	In the base case $q = 1$, meaning $D_kg$ is only a single-difference operator.
+	Since $s \geq 1$, replacing $g$ by its Taylor expansion inside the first order expected difference operator $\Ebb[D_kg(x)]$ yields
+	\begin{equation}
+	\label{eqn:leading_term_sobolev_compact_pf1}
+	\Ebb\Bigl[D_kg(x)\Bigr] = \sum_{1 \leq \abs{\alpha} < s} \Ebb_{\alpha}(x) \cdot g^{(\alpha)}(x)  + \sum_{\abs{\alpha} = s} \Bigl(\Ebb_{\alpha}\bigl[G_{\alpha}\bigr]\Bigr)(x)
+	\end{equation}
+	When $s = 1$ only the second term in the previous expression is non-zero, and we therefore begin by analyzing this term, obtaining that for each $\abs{\alpha} = s$,
+	\begin{align}
+	\biggl\|\Bigl(\Ebb_{\alpha}\bigl[G_{\alpha}\bigr]\Bigr)\biggr\|_{\Leb^2(\Xset_{r})}^2 & = \biggl\|\int (y - \cdot)^{\alpha} G_{\alpha}(\cdot,y) K_r(y,\cdot) p(y) \,dy\biggr\|_{\Leb^2(\Xset_{r})}^2 \nonumber \\
+	& \leq r^{2s} p_{\max}^2 K_{\max}^2  \biggl\|\int_{B(0,1)} G_{\alpha}(\cdot,zr + \cdot) \,dy\biggr\|_{\Leb^2(\Xset_{r})}^2 \nonumber \\
+	& \leq r^{2s} p_{\max}^2 K_{\max}^2 \nu_d \int_{B(0,1)} \Bigl\|G_{\alpha}(\cdot,zr + \cdot)\Bigr\|_{\Leb^2(\Xset_{r})}^2 \,dz \nonumber \\
+	& \leq r^{2s} p_{\max}^2 K_{\max}^2 \nu_d^2 \norm{g^{(\alpha)}}_{\Leb^2(\Xset)}^2 \nonumber \\
+	& \leq r^{2s} p_{\max}^2 K_{\max}^2 \nu_d^2 \norm{g}_{H^s(\Xset)}^2; \nonumber
+	\end{align}
+	hence~\eqref{eqn:leading_term_sobolev_compact_2} follows when $q = 1, s = 1$.
+
+	When $s \geq 2$ we must analyze $\Ebb_{\alpha}(x) \cdot g^{(\alpha)}(x)$, which we do by using the Taylor expansion of $p$.
+Since $B(x,r) \subset \Xset$, we recall that $\Ibb_{\alpha + \beta,x} = r^{\abs{\alpha} + \abs{\beta}}I_{\alpha,\beta}$; thus
+	\begin{align*}
+	\Ebb_{\alpha}(x) & = \int(y - x)^{\alpha} K_r(y,x) p(y) \,dy \\
+	& = \sum_{\abs{\beta} = 0}^{s - 2} p^{(\beta)}(x) \Ibb_{\alpha + \beta,x} + \sum_{\abs{\beta} = s - 1} \Bigl(\Ibb_{\alpha + \beta}\bigl[P_{\beta}\bigr]\Bigr)(x) \\
+	& = \sum_{\abs{\beta} = 0}^{s - 2} p^{(\beta)}(x) r^{\abs{\alpha} + \abs{\beta}} I_{\alpha + \beta} + \sum_{\abs{\beta} = s - 1} \Bigl(\Ibb_{\alpha + \beta}\bigl[P_{\beta}\bigr]\Bigr)(x).
+	\end{align*}
+	Replacing $\Ebb_{\alpha}(x)$ by this expansion in~\eqref{eqn:leading_term_sobolev_compact_pf1} gives
+	\begin{equation*}
+	\Ebb\Bigl[D_kg(x)\Bigr] = \sum_{\abs{\alpha} = 1}^{s - 1} \sum_{\abs{\beta} = 0}^{s - 2} r^{\abs{\alpha} + \abs{\beta}} I_{\alpha + \beta} g^{(\alpha)}(x) p^{(\beta)}(x)  + \sum_{\abs{\alpha} = 1}^{s} \sum_{\abs{\beta} = s - 1} g^{(\alpha)}(x) \Bigl( \Ibb_{\alpha + \beta}\bigl[P_{\beta}\bigr]\Bigr)(x)  + \sum_{\abs{\alpha} = s} \Bigl(\Ebb_{\alpha}\bigl[G_{\alpha}\bigr]\Bigr)(x)
+	\end{equation*}
+	We now divide the sum in the first term based on the size of $\abs{\alpha} + \abs{\beta}$.
+The critical fact is that $I_{\alpha + \beta} = 0$ when $\abs{\alpha} + \abs{\beta} = 1$.
+When $s = 2$ this leaves
+	\begin{equation*}
+	\Ebb\Bigl[D_kg(x)\Bigr] = \sum_{\abs{\alpha} = 1}^{s} \sum_{\abs{\beta} = s - 1} \Bigl(\Ibb_{\alpha + \beta}\bigl[P_{\beta}\bigr]\Bigr)(x) g^{(\alpha)}(x) + \sum_{\abs{\alpha} = s} \Bigl(\Ebb_{\alpha}\bigl[G_{\alpha}\bigr]\Bigr)(x).
+	\end{equation*}
+	We have already shown that the second term belongs to $\Leb^2(\Xset)$, and provided an appropriate upper bound on its norm.
+The first term is similarly upper bounded, since for each term inside the sum
+	\begin{equation*}
+	\norm{\Ibb_{\alpha + \beta}\bigl[P_{\beta}\bigr]g^{(\alpha)}}_{\Leb^2(\Xset)}^2 \leq r^{2(\abs{\alpha} + \abs{\beta})} p_{\max}^2 \norm{g^{(\alpha)}}_{\Leb^2(\Xset)}^2 \leq r^{2(\abs{\alpha} + \abs{\beta})} p_{\max}^2 \norm{g}_{H^s(\Xset)}^2 \nonumber
+	\end{equation*}
+	thus establishing~\eqref{eqn:leading_term_sobolev_compact_2} when $s = 2$. Otherwise when $s > 2$, we rearrange
+	\begin{equation*}
+	\sum_{\abs{\alpha} = 1}^{s} \sum_{\abs{\beta} = 0}^{s - 2} r^{\abs{\alpha} + \abs{\beta}} I_{\alpha + \beta} g^{(\alpha)}(x) p^{(\beta)}(x) = \sum_{\ell = 2}^{s - 1} r^{\ell} \Biggl\{\underbrace{\sum_{\abs{\alpha} + \abs{\beta} = \ell} I_{\alpha + \beta} g^{(\alpha)}(x) p^{(\beta)}(x)}_{:=g_{\ell,1}(x)}\Biggr\} + \sum_{\ell = s + 1}^{2s - 2} r^{\ell} \sum_{\abs{\alpha} + \abs{\beta} = \ell} I_{\alpha + \beta} g^{(\alpha)}(x) p^{(\beta)}(x).
+	\end{equation*}
+	and therefore
+	\begin{align*}
+	& \Ebb\Bigl[D_kg(x)\Bigr] - \sum_{\ell = 2}^{s - 1} r^{\ell} g_{\ell,1}(x) = \\ & ~~~~ \sum_{\ell = s + 1}^{2s - 2} r^{\ell} \sum_{\abs{\alpha} + \abs{\beta} = \ell} I_{\alpha + \beta} g^{(\alpha)}(x) p^{(\beta)}(x) +  \sum_{\abs{\alpha} = 1}^{s} \sum_{\abs{\beta} = s - 1} \Bigl(\Ibb_{\alpha + \beta}\bigl[P_{\beta}\bigr]\Bigr)(x) g^{(\alpha)}(x) + \sum_{\abs{\alpha} = s} \Bigl(\Ebb_{\alpha}\bigl[G_{\alpha}\bigr]\Bigr)(x)
+	\end{align*}
+	On the left hand side, taking $\ell = \abs{\alpha} + \abs{\beta}$, note that for each $\ell < s$ the function $g_{\ell,1} \in C_c^{s - \ell}(\Xset) \subset H_0^{s - \ell}(\Xset)$ and further
+	\begin{equation}
+	\label{eqn:leading_term_sobolev_compact_pf3}
+	\norm{g_{\ell,1}}_{H^{s - \ell}(\Xset)} \leq c p_{\max} \norm{g}_{H^s(\Xset)}.
+	\end{equation}
+
+	The right hand side consists of three terms, and we have already obtained sufficient estimates on the second and third term, so it remains to deal with the first term.
+We have that $g^{(\alpha)}\cdot p^{(\beta)} \in C_c^{0}(\Xset) \subset \Leb_0^2(\Xset)$ and
+	\begin{equation}
+	\label{eqn:leading_term_sobolev_compact_pf4}
+	\norm{g^{(\alpha)}p^{(\beta)}}_{\Leb^2(\Xset)} \leq p_{\max} \norm{g}_{H^s(\Xset)},
+	\end{equation}
+	establishing~\eqref{eqn:leading_term_sobolev_compact_1} when $s > 2$.
+
+	\paragraph{Induction step.}
+	We now assume that~\eqref{eqn:leading_term_sobolev_compact_2}-\eqref{eqn:leading_term_sobolev_compact_4} hold with respect to $g$ for all $k \in (n)^q$, and show the desired estimates on $\Ebb[D_jD_kg]$ for all $(kj) \in (n)^{q + 1}$.
+
+	We first consider the case where $s \leq 2q$. Then,
+	\begin{align*}
+	\norm{\Ebb\bigl[D_jD_kg\bigr]}_{\Leb^2(X_{(q + 1)r})}^2 \leq 2 p_{\max}^2 K_{\max}^2 \nu_d^2 \norm{D_kg}_{\Leb^2(X_{qr})}^2 \leq c r^{2s} \norm{g}_{H^s(\Xset)}
+	\end{align*}
+	where the final inequality follows by hypothesis, and gives the desired estimate.
+
+	Otherwise $s \geq 2q + 1$. We make use of the inductive hypothesis through the following three facts:
+	\begin{enumerate}
+		\item There exist functions $g_{2q,q}, \ldots, g_{s - 1,q}$ satisfying~\eqref{eqn:leading_term_sobolev_compact_4} such that
+		\begin{equation*}
+		\norm{\Ebb\bigl[D_kg\bigr] - \sum_{\ell = 2q}^{s - 1}r^{\ell}g_{\ell,q}}_{\Leb^2(X_{qr})} \leq c r^s \norm{g}_{H^s(\Xset)}
+		\end{equation*}
+		\item The functions $f_{\ell,q}$ belong to $H_0^{s - \ell}(\Xset)$. Thus by hypothesis, when $\ell = s - 1$ or $\ell = s - 2$,
+		\begin{equation*}
+		\norm{\Ebb\bigl[D_jg_{\ell,q}\bigr]}_{\Leb^2(\Xset_r)} \leq c r^s \norm{f}_{H^s(\Xset)}.
+		\end{equation*}
+		\item Otherwise if $s - \ell > 2$, there exist further functions $g_{\ell,m,q}$ for $m = 2,\ldots,s - \ell - 1$ such that
+		\begin{equation*}
+		\norm{\Ebb\bigl[D_jf\bigr] - \sum_{m = 2}^{s - \ell - 1}r^{m}g_{\ell,m,q}}_{\Leb^2(\Xset_r)} \leq  c r^s \norm{g}_{H^s(\Xset)}
+		\end{equation*}
+		The functions $g_{\ell,m,q} \in H_0^{s - (\ell + m)}(\Xset)$ additionally satisfy
+		\begin{equation*}
+		\norm{g_{\ell,m,q}}_{H^{s - (\ell + m)}(\Xset)} \leq c \norm{g_{\ell}}_{H^{s - \ell}(\Xset)} \leq c \norm{g}_{H^s(\Xset)}.
+		\end{equation*}
+	\end{enumerate}
+	Making use of the law of iterated expectation and the Fact 1, we have
+	\begin{align}
+	\Ebb\Bigl[D_jD_kg(x)\Bigr] & = \Ebb\biggl[\Bigl(\Ebb\bigl[D_kg(x_j)|x_j\bigr] - \Ebb\bigl[D_kg(x)\bigr]\Bigr)K_r(x_j,x)\biggr] \nonumber \\
+	& = \sum_{\ell = 2q}^{s - 1} r^{\ell} \Ebb\bigl[D_jg_{\ell}(x)\bigr] + \Ebb\biggl[\Bigl(\Ebb\bigl[D_kg\bigr](x_j) - \sum_{\ell = 2q}^{s - 1}r^{\ell} g_{\ell}(x_j)\Bigr)K_r(x_j,x)\biggr] + \Ebb\bigl[D_kg\bigr](x) - \sum_{\ell = 2q}^{s - 1}r^{\ell} g_{\ell}(x).
+\label{eqn:leading_term_sobolev_compact_pf6}
+	\end{align}
+	The above expansion consists of three terms. By Fact 1, the third term has bounded norm
+	\begin{equation*}
+	\norm{\Ebb\bigl[D_kg\bigr] - \sum_{\ell = 2q}^{s - 1}r^{\ell} g_{\ell}}_{\Leb^2(\Xset_{(q + 1)r})} \leq \norm{\Ebb\bigl[D_kg\bigr] - \sum_{\ell = 2q}^{s - 1}r^{\ell} g_{\ell}}_{\Leb^2(\Xset_{(q)r})} \leq c r^{s} \norm{g}_{H^s(\Xset)}
+	\end{equation*}
+	By Fact 1 and Lemma~\ref{lem:remainder_term_sobolev}, the same estimate holds with respect to the second term (up to constants).
+
+	When $s = (2q + 1)$ or $s = (2q + 2)$, by Fact 2
+	\begin{equation*}
+	\norm{ \sum_{\ell = 2q}^{s - 1} r^{\ell} \Ebb\bigl[D_jg_{\ell}\bigr]}_{\Leb^2(\Xset_r)} \leq c r^s \norm{g}_{H^s(\Xset)}
+	\end{equation*}
+	and the desired result~\eqref{eqn:leading_term_sobolev_compact_2} follows from the triangle inequality. Finally when $s > (2q + 2)$, by using Facts 2 and 3 we obtain
+	\begin{align*}
+	\norm{\sum_{\ell = 2q}^{s - 1} r^{\ell} \Ebb\bigl[D_jg_{\ell}\bigr] - \sum_{\ell = 2q}^{s - 3} \sum_{m = 2}^{s - \ell - 1} r^{\ell + m} g_{\ell,m,q}}_{\Leb^2(X_r)} \leq cr^s \norm{g}_{H^s(X)}
+	\end{align*}
+	Rewriting the double sum as a single sum over $\ell + m = 2q,\ldots,s - 1$ and plugging back in to~\eqref{eqn:leading_term_sobolev_compact_pf6} gives the desired result~\eqref{eqn:leading_term_sobolev_compact_4}.
 \end{proof}
+
+Here are a couple of Lemmas which helped us with the proof of Lemma~\ref{lem:leading_term_sobolev_compact}.
+
+\begin{lemma}
+	\label{lem:boundary_term_sobolev}
+	Let $\Xset \subset \Rd$ be a bounded open set with Lipschitz boundary. For any $g \in C_c^{\infty}(\Xset)$, we have that 
+	\begin{equation*}
+	\norm{g}_{\Leb^2(U_r)}^2 \leq c r^s \norm{g}_{H^s(\Xset)}
+	\end{equation*}
+	for all $r > 0$ sufficiently small.
+\end{lemma}
+\begin{proof}
+	Fix $x_0 \in \partial \Xset$, and let $Q_d(x_0,r)$ be the $d$-dimensional cube centered at $x_0$ of side length $r$. We will show that for all sufficiently small $r > 0$,
+	\begin{equation}
+	\norm{g}_{\Xset \cap \Leb^2(Q_d(x_0,r))} \leq c r^s \norm{g}_{H^s(Q_d(x_0,r))}
+	\end{equation}
+	The Lemma then follows by taking a finite covering of $\partial X$ -- possible since $X$ is assumed to be bounded -- in a similar manner to e.g.
+Theorem 18.1 of \textcolor{red}{(Leoni)} or Theorem 1 in 5.5 of \textcolor{red}{(Evans)}.
+
+	We begin by straightening the boundary.
+Since $\Xset$ has a Lipschitz boundary, for any $x_0 \in \partial X$ there exists a rigid motion $\Phi: \Rd \to \Rd$ with $T(x_0) = 0$, a Lipschitz continuous function $\gamma: \Reals^{d-1} \to \Reals^d$, and a radius $r_0' > 0$ such that, setting $y = \Phi(x)$ and writing $y = (y',y_d)$, we have that for all $r < r_0$,
+	\begin{equation*}
+	\Phi(\Xset \cap Q(x_0,r)) = \bigl\{y \in Q(0,r): y_d > \gamma(y')\bigr\}.
+	\end{equation*}
+	Fixing $0 < r < r_0/[\mathrm{Lip}(\gamma)\sqrt{d}]$, we have that $\gamma(y') > -r$ for all $y' \in Q_{d - 1}(x_0,r)$; writing $\wt{g}(y) = g(T^{-1}(y))$, taking a Taylor expansion of $\wt{g}(y)$ around $\wt{g}((y',\gamma(y')))$ thus yields
+	\begin{align*}
+	\wt{g}(y) & = \sum_{\ell = 1}^{s - 1} g^{(\ell e_d)}\bigl((y',\gamma(y'))\bigr) (y_d - \gamma(y'))^{\ell} + \int_{\gamma(y')}^{y_d} \wt{g}^{(se_d)}\bigl((y',he_d)\bigr) h^{s - 1}\,dh \\
+	& = \int_{\gamma(y')}^{y_d} \wt{g}^{(se_d)}\bigl((y',he_d)\bigr) h^{s - 1}\,dh
+	\end{align*}
+	where the second equality follows from the assumption $g \in C_c^{\infty}(\Xset)$. We now analyze the $\Leb^2$ norm over $Q_d(x_0,r)$,
+	\begin{align*}
+	\norm{g}_{\Xset \cap \Leb^2(Q_d(x_0,r))}^2 & = \norm{\wt{g}}_{\Leb^2(\Phi(\Xset \cap Q(0,r)))}^2 \\
+	& = \int_{Q_{d - 1}(0,r)} \int_{\gamma(y')}^{r} \biggl[\int_{\gamma(y')}^{y_d} \wt{g}^{(se_d)}\bigl((y',he_d)\bigr)h^{s - 1} \,dh\biggr]^2 \,dy_d \,dy' \\
+	& \overset{(i)}{\leq} (2r)^{2(s - 1)} \int_{Q_{d - 1}(0,r)} \int_{\gamma(y')}^{r} \bigl(y_d - \gamma(y')\bigr)^2 \biggl[\frac{1}{y_d - \gamma(y')}\int_{\gamma(y')}^{y_d} \wt{g}^{(se_d)}\bigl((y',he_d)\bigr) \,dh\biggr]^2 \,dy_d \,dy' \\
+	& \overset{(ii)}{\leq}  (2r)^{2(s - 1)} \int_{Q_{d - 1}(0,r)} \int_{\gamma(y')}^{r} \bigl(y_d - \gamma(y')\bigr) \int_{\gamma(y')}^{y_d} \biggl[\wt{g}^{(se_d)}\bigl((y',he_d)\bigr)\biggr]^2 \,dh \,dy_d \,dy' \\
+	& \overset{(iii)}{\leq}  (2r)^{2s - 1} \int_{Q_{d - 1}(0,r)} \int_{\gamma(y')}^{r} \int_{\gamma(y')}^{r} \biggl[\wt{g}^{(se_d)}\bigl((y',he_d)\bigr)\biggr]^2 \,dh \,dy_d \,dy' \\
+	& \leq  (2r)^{2s} \int_{Q_{d - 1}(0,r)}  \int_{\gamma(y')}^{r} \biggl[\wt{g}^{(se_d)}\bigl((y',he_d)\bigr)\biggr]^2 \,dh  \,dy' \\
+	& \overset{(iv)}{\leq} (2r)^{2s} \int_{Q(x_0,r)} \bigl[g^{(se_d)}(x)\bigr]^2 \,dx \leq (2r)^{2s} \norm{g}_{H^s(B(x_0,r))}^2
+	\end{align*} 
+	where $(i)$ follows since $0 < y_d - \gamma(y') < r - \gamma(y') < 2r$, $(ii)$ follows by Jensen's inequality, $(iii)$ follows since $y_d < r$, and $(iv)$ follows from a change of variables.
+This completes the proof of Lemma~\ref{lem:boundary_term_sobolev}.
+\end{proof}
+
+The following Lemma helps us deal with remainder terms in the Sobolev case.
+\begin{lemma}
+	\label{lem:remainder_term_sobolev}
+	Suppose $f \in \Leb^2(U_r)$ and $k \in (n)$. Then,
+	\begin{equation*}
+	\norm{\Ebb\bigl[D_kf\bigr]}_{\Leb^2(U)}, \norm{\Ebb\bigl[fK_r(x_k,\cdot)\bigr]}_{\Leb^2(U)} \leq c \norm{f}_{\Leb^2(U_r)}
+	\end{equation*}
+\end{lemma}
 

### 2020-02-07 notes/weekly_notes/2.7.20.NotesForWeek.tex
added proof of testing rate over W0

--- 2020-02-07 notes/graph_testing.tex
@@
 \newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
+\newcommand{\norm}[1]{\left \lVert #1 \right \rVert}
 \newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
@@
 \newcommand{\Rd}{\Reals^d}
+\newcommand{\Rdp}{\Reals_+^d}
 
@@
 \newcommand{\spec}{\mathrm{spec}}
+\newcommand{\dist}{\mathrm{dist}}
 
@@
 \begin{document}
-\title{Graph Testing}
+\title{Notes for Week 1/30/20 - 2/7/20}
 \author{Alden Green}
@@
 
-This document details the current status of the graph testing project.
-We divide by section based on the testing problem under consideration.
-At the end of each section, we list some areas we are interested in investigating.
-All proofs are left until the end.
-
-We establish some notation that we will use throughout.
-A graph $G = (V,W)$ consists of a collection of vertices $V = \{v_1,\ldots,v_n\}$, and a weight matrix $W \in \Reals^{\abs{V} \times \abs{V}}$ with $ij$th entry $W_{ij}$ encoding affinity between $v_i$ and $v_j$.
-We will sometimes write $W_G$ to make it clear which graph a weight matrix is associated with.
-When the weight matrix $W_G$ consists only of $0$s and $1$s, we will use the equivalent representation $G = (V,E(G))$ when convenient; here $E(G) = V \times V \cap \{(v_i,v_j): W_{ij} = 1\}$.
-
-We will focus our attention on neighborhood graphs, and operators associated with them.
-For a kernel $K:\Reals \to \Reals$ and radius $r > 0$, we define $G_{n,K}=(V,W)$ to be the neighborhood graph with vertices $V = X$, and weights $W_{ij} = K(\norm{x_i - x_j}/r)$.
-Of particular interest to us will be the \emph{random geometric graph} $G_{n,r}$, the neighborhood graph constructed using the uniform kernel $K(z) := \1\{z \leq 1\}$.
-The weight matrix $W(G_{n,r})$ consists only of $0$s and $1$s, and we will thus use the equivalent representation $G_{n,r} = (X,E(G_{n,r}))$.
-
-The primary operator we are interested in (at this point) is the graph Laplacian. Let $D_G$ be the diagonal degree
-matrix of a graph $G$, with entries $D_{uu} := \sum_{v \in V} W_{uv}$.
-The graph Laplacian is then $L_G = D_G - W_G$.
-Let $\Lambda(G)$ consist of the eigenvalues $0 = \lambda_1(G) \leq \lambda_2(G) \leq \cdots \leq \lambda_n(G)$ of $L_G$, with $v_k(G) = (v_{k,1}(G),\ldots,v_{k,n}(G)) \in \Reals^n$ denoting the eigenvector corresponding to the $k$th eigenvalue $\lambda_k(G)$.
-
-\section{Regression goodness-of-fit testing.}
-
-Suppose we observe samples $(y_i,x_i)$ for $i = 1,\ldots,n$, where conditional on $X$ the responses $Y = \{y_1,\ldots,y_n\}$ are assumed to follow the model
-\begin{equation}
-\label{eqn:regression_known_variance}
+Suppose we observe samples $(y_i,x_i)$ for $i = 1,\ldots,n$.
+Here $x_1,\ldots,x_n$ are random design points, sampled independently from a distribution $P$ with density $p$ supported on an open set $\Xset \subset \Rd$.
+The responses $Y = \{y_1,\ldots,y_n\}$ are then assumed to follow the model
+\begin{equation}
+\label{eqn:regression_random_design_known_variance}
 y_i = f(x_i) + \varepsilon_i, ~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
 \end{equation} 
-and our task is to distinguish
+Our task is to distinguish
 \begin{equation*}
@@
 
-\subsection{Test Statistics.}
-We list the test statistics we use for the regression testing problem.
-
-\paragraph{Eigenvector projection test statistic.}
-The graph Laplacian eigenvector projection test is a truncated-series test.
-Suppose we are given a graph $G = (X,W)$ defined on the sample points $X$.
-Letting $\kappa$ be some integer between $1$ and $n$, our test projects $Y$ onto the eigenvectors $v_1(G),\ldots,v_{\kappa}(G)$, and then takes the empirical norm of this projection to be the test statistic.
-Formally,
-\begin{equation*}
-T_{\mathrm{spec}}(G) := \frac{1}{n} \sum_{k = 1}^{\kappa} \Biggl(\sum_{i = 1}^{n} y_i v_{k,i}(G)\Biggr)^2
-\end{equation*}
-Then the test $\phi_{\spec}(G) := \1\{T_{\spec}(G) \geq \tau\}$ rejects the null hypothesis when the test statistic $T_{\mathrm{spec}}(G)$ is greater than a pre-specified cutoff $\tau$.
-
-\subsection{Random design.}
-In the random design case, we assume that each design point $x_i$ is sampled independently from some distribution $P$. Formally, our model for the samples is the following:
-\begin{equation}
-\label{eqn:regression_random_design_known_variance}
-x_1,\ldots,x_n \overset{\textrm{i.i.d}}{\sim} P,~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)~\textrm{and}~ \varepsilon \perp X,~~ y_i = f(x_i) + \varepsilon_i.
-\end{equation}
-
-In Theorem~\ref{thm:sobolev_testing_rate_order1} we show that under some typical regularity conditions on $P$, the neighborhood graph Laplacian eigenvector projection test $\phi_{\textrm{spec}}(G_{n,r})$ is a minimax optimal test over the Sobolev ball $H^1(\mathcal{X};R)$ for $d = 1,2,3$ and $\Xset = [0,1]^d$.
-Let $p_d$ equal $3/4$ when $d = 2$ and equal $1/d$ for $d \geq 3$.
-
-\begin{theorem}
-	\label{thm:sobolev_testing_rate_order1}
-	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}.
-Let $R > 0$ and $b \geq 1$ be fixed constants, and $d = 1,2$ or $3$.
-Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p(x)$ is bounded away from zero and infinity,
-	\begin{equation*}
-	0 < p_{\min} < p(x) < p_{\max} < \infty,~~ \textrm{for all $x \in \mathcal{X}$.}
-	\end{equation*}
-	and the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choices
-	\begin{equation*}
-	c \frac{(\log n)^{p_d}}{n^{1/d}} \leq r(n) \leq n^{-4/((4 + d)(2+d))}, ~\kappa = (nR)^{2d/(4 + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
-	\end{equation*}
-	for $c$ a constant that depends only on $\Xset, p_{\min}$ and $p_{\max}$.
-Then the following statements holds for every $n$ sufficiently large: there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
-	\begin{equation}
-	\label{eqn:sobolev_testing_rate_order1}
-	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4 + d)}\} \cdot n^{-4/(4 + d)}
-	\end{equation}
-	the worst-case risk is upper bounded
-	\begin{equation}
-	\label{eqn:sobolev_testing_rate_order1_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}(G_{n,r}); H^1(\mathcal{X};R)) \leq \frac{c_2}{b}.
-	\end{equation}
-\end{theorem}
-
-The testing problem in the random design case with known variance fundamentally changes in the low-smoothness regime, where the number of bounded derivatives is small compared to the dimension $d$.
-In the case where $f \in H^1(\Xset;R)$ is assumed to have one derivative which is bounded in $\Leb^2(\Xset)$ norm, this transition occurs at $d = 4$.
-We shall discuss testing in the low-smoothness regime shortly.
-First, let's consider the situation when we assume more derivatives worth of smoothness.
-
-Theorem~\ref{thm:sobolev_testing_rate_order1} establishes that a truncated series test using graph eigenvectors performs just as well, in a minimax sense, as a truncated series test using a classic Fourier series; for analysis of the latter test as well as derivation of the minimax rate of nonparametric goodness-of-fit testing over Sobolev classes see \textcolor{red}{(Ingster)}.
-Intuitively, we can understand this by thinking of the graph eigenvectors as estimates of eigenfunctions of a density-weighted continuum Laplacian operator.
-When the density is uniform, and the domain $\Xset$ is the unit cube, under appropriate boundary conditions these eigenfunctions are themselves elements of the Fourier basis.
-When the density is merely close to uniform, the eigenfunctions no longer belong to the Fourier basis; nevertheless, they share enough properties with the latter that the resulting minimax rate is unchanged.
-Theorem~\ref{thm:sobolev_testing_rate_order1} shows that the additional error due to using \textit{estimates} of these eigenfunctions does not fundamentally change the minimax rate.
-
-When we additionally believe $H^s(\Xset)$, we know (again from Ingster) that the minimax testing rate is $n^{-4s/(4s + d)}$.
-Now, however, our estimates of the eigenvectors no longer track the derivatives finely enough at the boundaries of $\Xset$ to establish the same minimax rate over all $H^s(\mathcal{X})$.
-However, if we assume $f$ and its derivatives are compactly supported on $\Xset$, we recover the usual rate.
-Theorem~\ref{thm:sobolev_testing_rate} presents our formal result, that $\phi_{\textrm{spec}}(G_{n,r})$ is a minimax optimal test over the Sobolev balls $H_0^s(\mathcal{X};R)$ whenever $4s > d$ and $\Xset = [0,1]^d$.
-
-\begin{theorem}
-	\label{thm:sobolev_testing_rate}
-	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}.
-Let $R > 0$, $b,s,d \geq 1$ be fixed constants, with $s$ and $d$ integers.
-Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p \in C^{s-1}(\Xset;p_{\max})$ for some $p_{\max} < \infty$, and further that $p(x)$ is bounded away from zero, i.e.
-there exists $p_{\min} > 0$ such that
-	\begin{equation*}
-	p_{\min} < p(x),~~ \textrm{for all $x \in \mathcal{X}$.}
-	\end{equation*}
-	Then the following statement holds for all $n$ sufficiently large: if the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choice
-	\begin{equation*}
-	n^{-1/(2(s-1) + d)} \leq r(n) \leq n^{-4/((4s + d)(2+d))}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
-	\end{equation*}
-	then there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
-	\begin{equation}
-	\label{eqn:sobolev_testing_rate}
-	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4s + d)}\} \cdot n^{-{4s}/(4s + d)}
-	\end{equation}
-	the worst-case risk is upper bounded
-	\begin{equation}
-	\label{eqn:sobolev_testing_rate_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; H_0^{s}(\mathcal{X};R)) \leq \frac{c_2}{b}.
-	\end{equation}
-\end{theorem}
-
-The restriction that $p$ be supported on the unit cube is mostly for convenience.
-If instead $p$ is supported on any compact set with Lipschitz boundary, the Theorem statement should hold (up to constants), with only a few modifications to our proofs.
-However, we do not work through the details.
-
-The restriction $4s > d$ is a fundamental consequence of the tightness of the Sobolev embedding theorem. When $4s \geq d$ the compact embedding
-\begin{equation*}
-H^{s}(\mathcal{X}) \subseteq \Leb^4(\mathcal{X}) 
-\end{equation*}
-does not hold (as it does when $d < 4s$). However, if we directly assume $f \in \mathcal{L}_d^4(\mathcal{X};R)$ (regardless of $d$) we obtain the following result.
-\begin{proposition}
-	\label{prop:L4_testing_rate}
-	If $f \in \Leb^4(\mathcal{X};R)$, there exists a constant $c$ such that if
-	\begin{equation}
-	\label{eqn:L4_testing_rate}
-	\epsilon^2 > b \cdot n^{-1/2}
-	\end{equation}
-	then the test
-	\begin{equation*}
-	\phi_{\mathrm{mean}} = \1\{\frac{1}{n}\sum_{i = 1}^{n} y_i^2 \geq 1\}
-	\end{equation*}
-	has worst-case risk
-	\begin{equation*}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; H^1(\mathcal{X};R)) \lesssim \frac{1}{b}. 
-	\end{equation*}
-\end{proposition}
-Note that when $4s > d$ the critical radius~\eqref{eqn:sobolev_testing_rate} is smaller than \eqref{eqn:L4_testing_rate}.
-
-\subsection{Analysis}
-
-To prove Theorem~\ref{thm:sobolev_testing_rate}, we show that there exists a high-probability set $E \subseteq \Xset^n$ such that conditional on $X \in E$, the test $\phi_{\spec}(G_{n,r})$ has small risk.
-Since $G_{n,r}$ is a function only of $X$ and not of $Y$, this amounts to reasoning about the behavior of the test $\phi_{\spec}$ over a fixed graph $G = (X,E)$, where we observe
-\begin{equation}
-y_i = \beta_i + \varepsilon_i,~~\varepsilon_i \sim \mathcal{N}(0,1)
-\end{equation}
-for some fixed $\beta \in \Reals^n$.
-In Lemma~\ref{lem:fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}(G)$.
-Our bound on the Type II error will be stated as a function of $\beta^T L^s \beta$--a measure of the smoothness the signal $\beta$ displays over the graph $G$--as well as the $\kappa$th eigenvalue $\lambda_{\kappa}$.
+\paragraph{Test statistics.}
+
+For a graph $G$ with Laplacian $L_G = D_G - A_G$, we arrange the eigenvalues in ascending order and denote them $\lambda_1(G) \leq \ldots \leq \lambda_n(G)$.
+We let $v_k(G)$ be the eigenvector associated with the $k$th eigenvalue $\lambda_k(G)$.
+
+We define the graph Laplacian eigenvector projection test to be the magnitude of the projection of the responses $y$ onto the first $\kappa$ graph Laplacian eigenvectors, 
+\begin{equation*}
+\phi_{\spec}(G_{n,r}) := \1\bigl\{T_{\textrm{spec}}(G_{n,r}) \geq \tau\bigr\},~~ T_{\textrm{spec}}(G_{n,r}) := \frac{1}{n}\sum_{k = 1}^{\kappa} \biggl(\sum_{i = 1}^{n} v_{k,i}(G_{n,r}) y_i\biggr)^2.
+\end{equation*}
+
+The action of the graph Laplacian $L_{n,r}f(x)$ (we use the abbreviation $L_{n,r} := L_{G_{n,r}}$ to avoid the double subscripts) can be viewed as approximating the action of a continuum weighted Laplacian operator $\varDelta_Pf(x)$.
+The latter operator is a 2nd order differential operator; however the graph Laplacian does a poor job of tracking the 2nd derivative of $f$ when $x$ is sufficiently close to the boundary of the domain $\Xset$.
+The problem is only exacerbated when we consider the iterated graph Laplacian operator $L_{n,r}^sf(x)$, and view it as an estimate of the continuum weighted Laplacian operator $\varDelta_P^sf(x)$.
+To rectify the situation, we will assume that $f \in C_c^{s}(\Xset)$, so that tracking derivatives near the boundary is no longer an issue.
+We let $\Xset_r := \{x \in \Xset: \dist(x,\Xset) > r\}$ denote the remainder of $\Xset$ once a tube of radius $r$ around the boundary has been removed.
 
 \begin{lemma}
-	\label{lem:fixed_graph_testing}
-	Let $1 \leq \kappa \leq n$ be an integer.
-	\begin{enumerate}
-		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}$ is upper bounded
-		\begin{equation}
-		\label{eqn:graph_spectral_type_I_error}
-		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{1}{b^2}.
-		\end{equation}
-		\item \textbf{Type II error:} For any $b \geq 1$ and $\beta$ such that
-		\begin{equation}
-		\label{eqn:fixed_graph_testing_critical_radius}
-		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{\beta^T L^s \beta}{n\lambda_{\kappa}^s}
-		\end{equation}
-		the Type II error of $\phi_{\spec}$ is upper bounded,
-		\begin{equation}
-		\label{eqn:graph_spectral_type_II_error}
-		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{3}{b}.
-		\end{equation}
-	\end{enumerate}
+	\label{lem:leading_term_holder_compact}
+	Fix integers $s \geq 0$ and $q \geq 1$, and index vector $k \in (n)^q$.
+Suppose that $f \in C_c^s(\Xset)$, that $p \in C^0(\Xset;p_{\max})$, and additionally that $p \in C^{s-1}(\Xset;p_{\max})$ if $s \geq 2$.
+	\begin{itemize}
+		\item If $x \in \Xset_{qr}$ and $2q < s$, then there exist functions $f_{\ell,q} \in C_c^{s - \ell}(\Xset)$ which additionally satisfy
+		\begin{equation}
+		\label{eqn:leading_term_holder_compact_0}
+		\norm{f_{\ell,q}}_{C^{s - \ell}(\Xset)} \leq c \norm{f}_{C^s(\Xset)},
+		\end{equation}
+		for each $\ell = 2q,\ldots,s-1$, such that 
+		\begin{equation}
+		\label{eqn:leading_term_holder_compact_1}
+		\abs{\Ebb\Bigl[D_kf(x)\Bigr] -
+		\sum_{\ell = 2q}^{s - 1} f_{\ell,q}(x) r^{\ell}} \leq c r^s \norm{f}_{C^s(\Xset)} ~~\textrm{if $2q < s$,}
+		\end{equation},
+		\item Otherwise if $2q \geq s$ or $x \in \Xset \setminus \Xset_{qr}$,
+		\begin{equation}
+		\label{eqn:leading_term_holder_compact_2}
+		\biggl|\Ebb\Bigl[D_kf(x)\Bigr]\biggr| \leq cr^s\norm{f}_{C^s(\Xset)}.
+		\end{equation}
+	\end{itemize}
 \end{lemma}
-
-\subsection{Areas to Investigate}
-
-\begin{enumerate}[(i)]
-	\item When $d \geq 4$, is it true that there does not exist a uniformly consistent
-	test over the Sobolev ball $W_d^{1,2}(\mathcal{X};R)$?
-	\item Assume the distribution $P$ is supported on a manifold $\mathcal{M}$ of intrinsic dimension $s < d$.
-Does $\mathrm{\phi_{\mathrm{spec}}}$ display adaptivity to the intrinsic dimension of $\mathcal{M}$?
-	\item Assume that $f$ belongs to the Holder space $C_d^s(\mathcal{X})$.
-Moreover, suppose that instead of observing ${y_i}$ according to the regression testing model \eqref{eqn:regression_known_variance}, we observe
-	\begin{equation*}
-	y_i = f(x_i) + \sigma \varepsilon_i, ~\varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
-	\end{equation*}
-	where $\sigma > 0$ is unknown.
-When $d \geq 4$, what are the minimax regression testing rates over $C_d^1(\mathcal{X};R)$?
-Is the test $\phi_{\mathrm{spec}}$ minimax optimal, when the tuning parameters $r$ and $\kappa$ are appropriately chosen?
-\end{enumerate}
-
-\section{Two-sample density testing.}
-In the two-sample density testing problem, we observe independent samples $Z = z_1,\ldots,z_N \sim P$ and $Y = y_1,\ldots,y_M \sim Q$, where $P$ and $Q$ are distributions over $\Reals^d$ with densities $p$ and $q$, respectively, and $N \sim \textrm{Bin}(n,1/2)$.
-Our goal is to distinguish the hypotheses
-\begin{equation*}
-\mathbf{H}_0: P = Q \quad \textrm{vs} \quad \mathbf{H}_{\textrm{a}}: P \neq Q
-\end{equation*}
-and we again evaluate our performance using worst-case risk; letting $\phi:\Reals^{N + M} \to \{0,1\}$, 
-\begin{equation*}
-\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \inf_{p \in \mathcal{H}}\Ebb_{p,p}(\phi) + \sup_{\substack{p,q \in \mathcal{H} \\ \norm{p - q}_{\Leb^2} \geq \epsilon}} \Ebb_{p,q}(1 - \phi).
-\end{equation*}
-
-\subsection{Test statistics.}
-
-We suggest several two-sample test statistics.
-
-\paragraph{Eigenvector projection test statistic.}
-
-It is straightforward to adapt the test statistic $T_{\mathrm{spec}}$ to the two-sample testing problem.
-Concatenate the samples $Z$ and $Y$ in $X = (z_1,\ldots,z_N,y_1,\ldots,y_M)$, and define $T_{\mathrm{spec}}^{(2)}$ to be
-\begin{equation}
-\label{eqn:graph_spectral_projections_2}
-T_{\mathrm{spec}}^{(2)} := \frac{1}{n} \sum_{k = 0}^{\kappa} \left(\sum_{i = 1}^{n} v_i a_i\right)^2, ~~\textrm{where}~~ a = (\underbrace{N^{-1},\ldots,{N^{-1}}}_{\textrm{length } N},\underbrace{-M^{-1},\ldots,-M^{-1}}_{\textrm{length } M})
-\end{equation}
-
-For convenience, we state our following two test statistics with respect to the empirical norm $\norm{\theta}_n = n^{-1/2}\norm{\theta}_2$ for $\theta \in \Reals^n$.
-They will each depend on a tuning parameter $\lambda > 0$.
-\paragraph{Graph Sobolev IPM.}
-Letting $C_n := nr^{(d + 2)/2}$ and
-\begin{equation*}
-\Theta_{1,2} := \{\theta \in \Reals^n:~ C_n^{-1} \norm{B\theta}_2 \leq 1\} 
-\end{equation*}
-we define the \emph{graph Sobolev IPM} to be
-\begin{equation}
-\label{eqn:sobolev_IPM}
-T_{\textrm{sob}} := \sup_{\substack{\theta \in \Theta_{1,2} \\ \lambda \norm{\theta}_n \leq 1}} \abs{\frac{1}{n}\sum_{i = 1}^{n} a_i \theta_i}. 
-\end{equation}
-
-\paragraph{Graph Total Variation IPM.}
-Letting $C_n' := n^{2}r^{(d + 1)}$ and 
-\begin{equation*}
-\Theta_{1,1} := \{\theta \in \Reals^n:~ (C_n')^{-1} \norm{B\theta}_1 \leq 1\}
-\end{equation*}
-we define the \emph{graph Total Variation} IPM to be
-\begin{equation}
-\label{eqn:total_variation_IPM}
-T_{\mathrm{TV}} := \sup_{\substack{\theta \in \Theta_{1,1}, \\ \lambda \norm{\theta}_n \leq 1} } \abs{\frac{1}{n}\sum_{i = 1}^{n} a_i \theta_i}, \quad
-\end{equation}
-
-\subsection{Current results.}
-
-In Theorem~\ref{thm:twosample_sobolev_testing_rate} we show that under some typical regularity conditions on $P$, the test $\phi_{\textrm{spec}}^{(2)} := \1\{T_{\mathrm{spec}}^{(2)} \geq \tau\}$ is, up to log factors, a minimax optimal test over the Sobolev ball $H^1(\mathcal{X};R)$ when $d = 1$.
-
-\begin{theorem}
-	\label{thm:twosample_sobolev_testing_rate}
-	Let $b \geq 1$ and $a > 0$ be fixed constants, and let $d = 1$.
-Suppose that $\mu = (P + Q)/2$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]$ with density functions $\rho$ bounded above and below by constants, i.e
-	\begin{equation*}
-	0 < \rho_{\min} < \rho(x) < \rho_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
-	\end{equation*}
-	Then the following statement holds: if the test $\phi_{\spec}^{(2)}$ is performed with parameter choices 
-	\begin{equation*}
-	r = \log^a n \cdot \left(\frac{\log n}{n}\right), ~\kappa = n^{2/5}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
-	\end{equation*}
-	then there exists a constant $c$ which may depend on $R,p_{\max},q_{\max},b$ and $a$ but is independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
-	\begin{equation}
-	\label{eqn:twosample_sobolev_testing_rate}
-	\epsilon^2 \geq c \cdot b \cdot n^{-4/5} (\log n)^{h(a,1)}
-	\end{equation}
-	the worst-case risk is upper bounded
-	\begin{equation}
-	\label{eqn:twosample_sobolev_testing_rate_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; \mathcal{W}^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}.
-	\end{equation}
-\end{theorem}
-
-We prove Theorem~\ref{thm:twosample_sobolev_testing_rate} by relating the density testing problem to a regression testing problem with a certain type of structured noise, and then proceeding along similar lines to the proof of Theorem~\ref{thm:sobolev_testing_rate}.
-To pursue this strategy, we require the eigenvectors to satisfy a certain type of incoherence condition; this is in constrast to the regression testing problem with known variance, where we did not require the eigenvectors to be smooth in any sense.
-
-\subsection{Areas to Investigate}
-
-\begin{enumerate}[(i)]
-	\item When $1 < d < 4$, is the test $\phi_{\spec}^{(2)}$ minimax optimal?
-	\item Are either or both of the test statistics \eqref{eqn:graph_spectral_projections_2}-\eqref{eqn:sobolev_IPM} minimax optimal over $C_d^1(\mathcal{X};R)$ for all values of $d$?
-	\item Are either or both of the test statistics \eqref{eqn:graph_spectral_projections_2}-\eqref{eqn:sobolev_IPM} minimax optimal over $W_d^{1,2}(\mathcal{X};R)$ when $d \leq 4$?
-	\item Is the test statistic \eqref{eqn:graph_spectral_projections_2}, computed over a graph with suitable choice of kernel $K$, minimax optimal over higher order derivative classes $C_d^s$ and $H^s$?
-	\item Modify the test statistic \eqref{eqn:sobolev_IPM} by replacing the function class $\Theta_{1,2}$ with
-	\begin{equation}
-	\Theta_{s,2} := \{\theta \in \Reals^n:~ C_n^{-1} \norm{B^{(s)}\theta}_2 \leq 1\} 
-	\end{equation}
-	Is the modified test statistic, computed over a graph with suitable choice of kernel $K$, minimax optimal over higher order derivative classes $C_d^s$ and $H^s$?
-	\item What is the minimax testing rate over $BV_d^{1}(\mathcal{X};R)$? Does it exhibit a phase transition analogous to the minimax estimation rate over bounded variation spaces?
-	\item Is the test statistic \eqref{eqn:total_variation_IPM} minimax optimal over $BV_d^{1}(\mathcal{X};R)$?
-	\item Is the test statistic \eqref{eqn:total_variation_IPM} minimax optimal over Sobolev and Holder function classes?
-\end{enumerate}
-
-\section{Definitions and Notation}
-
-Here we collect definitions of some common function spaces and graph operators.
-
-\subsection{Function Spaces}
-
-\paragraph{Lebesgue spaces.}
-
-We say a Borel measurable function $f: \mathcal{X} \to \Reals$ is in the space $\mathcal{L}^p(\mathcal{X})$ for $1 \leq p < \infty$ if 
-$$\norm{f}_{\mathcal{L}^p(\mathcal{X})} := \int_{\mathcal{X}} \abs{f(x)}^p \,dx < \infty$$
-and we let 
-\begin{equation*}
-\mathcal{L}^p(\mathcal{X};R) = \set{f \in \mathcal{L}^p(\mathcal{X}): \norm{f}_{\mathcal{L}^p(\mathcal{X})} < R}
-\end{equation*}
-be a ball in the Lebesgue space.
-
-\paragraph{Holder spaces.}
-
-For a given $s > 0$, let $\ell = \floor{s}$ be the largest integer strictly less than $s$. Then the $s$th Holder norm is given by
-\begin{equation*}
-\norm{f}_{C^{s}(\mathcal{X})} := \sum_{\abs{\alpha} < s} \norm{D^{\alpha}f}_{\infty} + \sum_{\abs{\alpha} = \ell} \sup_{x,y \in \mathcal{X}} \frac{\abs{D^{\alpha}f(y) - D^{\alpha}f(x)}}{\norm{x - y}_2^{s - \ell}}
-\end{equation*}
-and the $s$th Holder space $C^{s}(\mathcal{X})$ consists of all functions which are $\ell$ times continuously differentiable with finite Holder norm $\norm{\cdot}_{C^{s}(\mathcal{X})}$.
-Denote the Holder ball by $C_d^{s}(\mathcal{X},R) = \set{f \in C^{s}(\mathcal{X}): \norm{f}_{C_d^{s}(\mathcal{X})} \leq R}$.
-Let $C_c^{s}(\Xset)$ consist of those functions in $C^s(\Xset)$ which are compactly supported within $\Xset$; formally $f \in C_c^{s}(\Xset)$ if and only if $f \in C^s(\Xset)$ and there exists an open set $V \subset \ol{V} \subset \Xset$ such that $\mathrm{supp}(f) \subset V$.
-
-\paragraph{Sobolev spaces.}
-
-For a given $s$ and $p > 0$, the Sobolev space $W^{s,p}(\mathcal{X})$ consists of all functions $f \in \Leb^2(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^2(\mathcal{X})$.
-The Sobolev $\{s,p\}$ norm is then
-\begin{equation*}
-\norm{f}_{W^{s,p}(\mathcal{X})}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^p \,dx
-\end{equation*}
-and for a given $R > 0$, the corresponding ball is $W^{s,p}(\Xset; R) = \set{f: \norm{f}_{H^s(\Xset)} \leq R}$.
-In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space; we adopt the usual convention of writing $H^s(\Xset) = H^s(\Xset)$.
-
-We write $W_0^{s,p}(\Xset)$ for the completion of $C_c^{\infty}(\Xset)$, and let $H_0(\Xset) = W_0^{s,2}(\Xset)$.
-
-\paragraph{Bounded Variation spaces.}
-
-For a function $f \in L^1(\mathcal{X})$ the \emph{total variation} semi-norm of $f$ is
-\begin{equation*}
-TV(f;\mathcal{X}) := \sup \left\{ \int_{\mathcal{X}} f \, \Xsetive \, \psi \,dx : \psi \in C_c^1(\mathcal{X}; \Reals^d), \abs{\psi} \leq 1 \right\};
-\end{equation*}
-and we write $BV_d(\mathcal{X})$ for the subset of functions $f \in L^1(\mathcal{X})$ which have bounded norm
-\begin{equation*}
-\norm{f}_{BV_d(\mathcal{X})} := \norm{f}_{\infty} + TV(f;\mathcal{X}).
-\end{equation*}
-For a given $R > 0$, the corresponding ball is $BV_d^{1}(\mathcal{X};R) = \set{f: \norm{f}_{BV_d(\mathcal{X})} \leq R}$.
-
-\subsection{Graph Operators}
-Let $s \geq 1$ be an integer. The $s$th-order difference operator on $G_{n,r}$, denoted $B^{(s)}$, is defined by
-\begin{equation*}
-B^{(s)} :=
-\begin{cases}
-L^{s/2},& ~~ s \textrm{ even} \\
-BL^{(s - 1)/2},& ~~ s \textrm{ odd.}
-\end{cases}
-\end{equation*}
-
-\subsection{Kernels}
-We say a kernel function $K(\cdot)$ is a $2$nd order kernel if $K$ is compactly supported on $B(0,1)$, uniformly upper bound $\abs{K(x)} \leq K_{\max} < \infty$ for all $x \in B(0,1)$, and
-\begin{equation*}
-\int K(x) \,dx = \nu_d,~~ \int x K(x) \,dx = 0.
-\end{equation*}
-Note that the uniform kernel, defined as $K(x) = \1\{\norm{x} \leq 1\}$, is a $2$nd order kernel.
-
-\subsection{Notation}
-
-We will treat $f$ interchangeably as a function $f:\Rd \to \Reals$, and as a vector $f = (f(x_1),\ldots,f(x_n))$.
-More generally, for $U \subset V$ and $f:V \to \Reals$, the action of $f$ on $U$ is naturally defined via the restriction operator $R_U$; we will forego additional notation and simply treat $f$ as a function on $U$, defined as $f(x) = R_Uf(x)$ for $x \in U$.
-It should always be clear from context how we are using $f$.
-
-For sequences $(a_n), (b_n)$, we say $a_n = O(b_n)$ if there exists a constant $c$ such that $a_n \leq cb_n$ for all $c$. We use the notation $a \wedge b$ for the minimum of $a$ and $b$.
-
-We use the notation $\Lambda(H)$ to denote the spectrum of a matrix $H$, and $\lambda_k(H)$ to denote the $k$th smallest eigenvalue of $H$.
-
-\section{Proofs}
-
-\subsection{Proof of Lemma~\ref{lem:fixed_graph_testing}}
-
-In this proof, we will drop all notational dependence on the graph $G$ for ease of reading.
-
-To prove Lemma~\ref{lem:fixed_graph_testing} we will first compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}$, and then use Chebyshev's inequality to show \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.
-
-\paragraph{Mean of $T_{\mathrm{spec}}$:} Using the notation $\dotp{v}{w} = \sum_{i = 1}^{n} v_iw_i$, we have
-\begin{align*}
-\Ebb(T_{\spec}) & = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{\beta}{v_k}^2 + \Ebb\bigl( \dotp{\varepsilon}{v_k}^2 + 2 \dotp{\varepsilon}{v_k} \dotp{\beta}{v_k}\bigr)\right) \\
-& = \frac{\kappa}{n} + \frac{1}{n}\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2.
-\end{align*}
-When $\beta = 0$, this equals $\kappa/n$. Otherwise, we have the following lower bound:
-\begin{align*}
-\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2 & = \norm{\beta}_2^2 - \sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \\
-& \geq \norm{\beta}_2^2 - \frac{1}{\lambda_{\kappa}^s}\sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \lambda_k^s \\
-& \geq \norm{\beta}_2^2 - \frac{\beta^T L^s \beta}{\lambda_{\kappa}^s},
-\end{align*}
-and therefore $\Ebb(T_{\spec}) \geq \kappa/n + n^{-1}(\norm{\beta}_2^2 - \beta^T L^s \beta/\lambda_{\kappa}^s)$.
-
-\paragraph{Variance of $T_{\mathrm{spec}}$:}
-We write $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvector $v_k$ as its $k$th column.
-\begin{align*}
-\Var(T_{\spec}) & = \frac{1}{n^2} \Var(y^T V_{\kappa} V_{\kappa}^T y) \\
-& = \frac{1}{n^2} \Var((\beta + \varepsilon)^T V_{\kappa} V_{\kappa}^T (\beta + \varepsilon)) \\
-& = \frac{1}{n^2} \Var(2 \beta^T V_{\kappa} V_{\kappa}^T \varepsilon + \varepsilon^T V_{\kappa} V_{\kappa}^T \varepsilon) \\
-& \leq \frac{1}{n^2}(4 \beta^T V_{\kappa} V_{\kappa}^T \beta + 2\kappa)
-\end{align*}
-where the last inequality follows from standard properties of the Gaussian distribution.
-We now move on to showing the desired inequalities \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.
-
-\paragraph{Proof of~\eqref{eqn:graph_spectral_type_I_error}:} By Chebyshev's inequality,
-\begin{align*}
-\Pbb_{\beta = 0}\bigl(T_{\spec} \geq \frac{\kappa}{n} + t(b)\bigr)
-& \leq \Pbb_{\beta = 0}\bigl(\abs{T_{\spec} - \frac{\kappa}{n}} \geq t(b)\bigr) \\
-& \leq \frac{\Var_{\beta = 0}(T_{\spec})}{t(b)^2} = \frac{1}{b^2}.
-\end{align*}
-
-\paragraph{Proof of~\eqref{eqn:graph_spectral_type_II_error}:} For simplicity, we introduce the notation
-\begin{equation*}
-\Delta = \frac{\norm{\beta}_2^2}{n} - \frac{\beta^T L^s \beta}{n\lambda_{\kappa}^s}.
-\end{equation*}
-Assumption~\eqref{eqn:fixed_graph_testing_critical_radius} implies $\Delta \geq 2 t(b)$. Then another application of Chebyshev's inequality gives us
-\begin{align*}
-\Pbb_{\beta}\bigl(T_{\spec} \leq \frac{\kappa}{n} + t(b)\bigr) & = \Pbb_{\beta}\bigl(T_{\spec} - \Ebb_{\beta}(T_{\spec}) \leq t(b) - \Delta \bigr) \\
-& \leq \Pbb_{\beta}\bigl(\abs{T_{\spec} - \Ebb_{\beta}(T_{\spec})} \leq \Delta - t(b) \bigr) \tag{since $\Delta \geq t(b)$}	\\
-& \leq \frac{\Var_{\beta}(T_{\spec})}{(\Delta - t(b))^2} \\
-& \leq 4\frac{\Var_{\beta}(T_{\spec})}{\Delta^2} \tag{since $\Delta \geq 2t(b)$} \\
-& \leq 4\frac{2\kappa/n^2 + 4\beta^T V_{\kappa} V_{\kappa}^T \beta /n^2}{\Delta^2}.
-\end{align*}
-We handle each summand in the numerator separately. For the first term, since $\Delta \geq 2 t(b)$, we have
-\begin{equation}
-\label{eqn:spectral_type_II_error_pf1}
-\frac{2\kappa}{n^2\Delta^2} \leq \frac{1}{2b^2}.
-\end{equation}
-
-For the second term, noting that $\Delta = \beta^T V_{\kappa} V_{\kappa}^T \beta/n$, we have
-\begin{align}
-\frac{\beta^T V_{\kappa} V_{\kappa}^T \beta/n^2}{\Delta^2} & = \frac{1}{n\Delta} \nonumber \\
-& \leq \frac{1}{2nt(b)} \nonumber \\
-& = \frac{1}{2b\sqrt{2\kappa}}, \label{eqn:spectral_type_II_error_pf2}
-\end{align}
-and combining~\eqref{eqn:spectral_type_II_error_pf1} and~\eqref{eqn:spectral_type_II_error_pf2} yields~\eqref{eqn:graph_spectral_type_II_error}.
-
-\subsection{Proof of Theorem~\ref{thm:sobolev_testing_rate_order1}}
-
-We want to apply Lemma~\ref{lem:fixed_graph_testing} to the case where $G = G_{n,r}$.
-In order to do this, we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$.
-In particular, we will show that the following statements each hold with probability at least $1 - c/b$ for sufficiently large $n$ (Here and in what follows, $c$ denotes a constant which is fixed in $n$ and $f$ and does not depend on $b$, but may depend on other fixed quantities such as $d$, $s$, etc.):
-\begin{enumerate}[label=(E\arabic*)]
-	\item 
-	\label{event:discrete_sobolev_norm_order1}
-	\textbf{Graph Sobolev norm:} For any $n^{-1/(2 + d)}\leq r \leq 1$,
-	\begin{equation}
-	\label{eqn:continuous_to_discrete_sobolev_norm_order1}
-	f^T L f \leq c \cdot b \cdot \norm{f}_{H^1(\Xset)}^2 \cdot n^{2} r^{d + 2} 
-	\end{equation}
-	\item 
-	\label{event:eigenvalue_tail_decay}
-	\textbf{Eigenvalue tail bound:} Let $t = n^{1/d}$, and let
-	\begin{equation*}
-	\overline{X} := \Bigl\{\frac{1}{t}(j_1,\ldots,j_d): k \in [t]^d\Bigr\}
-	\end{equation*} 
-	consist of $n$ evenly spaced grid points on $[0,1]^d$. The transportation distance between $X$ and $\ol{X}$ is
-	\begin{equation*}
-	d_{\infty}(X,\overline{X}) := \inf_{\pi} \sup\Bigl\{\norm{\pi(\ol{x}_i) - \ol{x}_i}: \ol{x}_i \in \ol{X}\Bigr\}
-	\end{equation*}
-	where the infimum is taken over all bijections $\pi: \ol{X} \to X$. 
-	There exists a constant $c$ such that for any $\kappa = 1,\ldots,n$ if $\max\{ 3 d_{\infty}(X,\ol{X}),n^{-1/d}\} \leq r \leq \kappa^{-2/(d(2 +d))}$, 
-	\begin{equation}
-	\label{eqn:eigenvalue_tail_bound}
-	\lambda_{\kappa} \geq c \cdot n r^{d + 2} \kappa^{2/d}
-	\end{equation}
-	Theorem~\ref{thm:slepcev_transport_distance} (Theorem 1.1 of \textcolor{red}{(Garcia Trillos and Slepcev)}) establishes that there exists a constant $c$ such that 
-	\begin{equation*}
-	d_{\infty}(X,\ol{X}) \leq c\biggl(\frac{\log n}{n}\biggr)^{1/d}
-	\end{equation*}
-	with probability $1  - O(1/n)$.
-Therefore in particular, the inequality~\eqref{eqn:eigenvalue_tail_bound} is satisfied for sufficiently large $n$ when $\kappa = n^{2d/(4 + d)}$ and $c(\log n/n)^{1/d} \leq r \leq n^{-4/((2+d)(4s + d))}$.
-	\item 
-	\label{event:l2_norm}
-	\textbf{Empirical norm of $f$:} There exists a constant $c_1$ such that if $\norm{f}_{\Leb^2(\Xset)} \geq c_1 \cdot b \cdot n^{-2/(4 + d)}$, and $d = 1,2$ or $3$,
-	\begin{equation}
-	\label{eqn:l2_to_empirical_norm}
-	\norm{f}_n^2 \geq \frac{1}{b} \cdot \norm{f}_{\Leb^2}^2
-	\end{equation}
-\end{enumerate}
-
-\subsubsection{Proof of \eqref{eqn:continuous_to_discrete_sobolev_norm}}
-
-The proof of~\eqref{eqn:continuous_to_discrete_sobolev_norm} follows from Lemma~\ref{lem:expected_first_order_seminorm}--which upper bounds the expected first order graph Sobolev seminorm-- and then an application of Markov's inequality.
-
-\subsubsection{Proof of \eqref{eqn:eigenvalue_tail_bound}}
-\label{subsubsec:eigenvalue_tail_bound_pf}
-
-We prove~\eqref{eqn:eigenvalue_tail_bound} by comparing $G_{n,r}$ to the tensor product of a $d$-dimensional lattice and a complete graph.
-The latter is a highly structured graph with known eigenvalues, which as we will see are sufficiently lower bounded for our purposes.
-
-Let $\wt{r} = r/(3(\sqrt{d} + 1)), M = (1/\wt{r})^d$ and $N = n\wt{r}^d$; assume without loss of generality that $M$ and $N$ are integers. Additionally for $m = M^{1/d}$ define
-\begin{equation*}
-\overline{Z} = \set{\frac{1}{m}(j_1,\ldots,j_d): j \in [m]^d}
-\end{equation*}
-to be the $M$ evenly spaced grid points over $[0,1]^d$.
-For a given $\overline{z}_j \in \overline{Z}$, we write $Q(z_j) = m^{-1}[j_1 - 1,j_1] \times \cdots \times m^{-1}[j_d - 1,j_d]$ for the cube of side length $1/m$ with $z_j$ at one corner.
-
-Consider the graph $H = (\overline{X}, E_H)$, where $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$ if
-\begin{equation*}
-\textrm{there exists}~\ol{z}_i, \ol{z}_j \in \ol{Z}~\textrm{such that}~\ol{x}_k \in Q(\ol{z}_i),~ \ol{x}_\ell \in Q(\ol{z}_j),~\textrm{and}~\norm{i - j}_1 \leq 1.
-\end{equation*}
-On the one hand $H \cong \ol{G}^M_d \otimes K_N$ where $\ol{G}^M_d$ is the $d$-dimensional lattice on $M$ nodes, and $K_N$ is the complete graph on $N$ nodes.
-On the other hand, we now show that when $\max\{3d_{\infty}(X,\ol{X}), n^{-1/d}\} \leq r$ then $G_{n,r} \succeq H$ as a result of the triangle inequality.
-If $(\ol{x}_k, \ol{x}_{\ell}) \in E(H)$, by definition there exist $\ol{z}_i, \ol{z}_j$ connected in $\ol{G}_d^M$ such that $\ol{x}_k \in Q(\ol{z}_i)$ and $\ol{x}_{\ell} \in Q(\ol{z}_j)$.
-This implies that $\ol{x}_k$ and $\ol{x}_{\ell}$ must themselves be close together, since
-\begin{equation*}
-\norm{\ol{x}_k - \ol{x}_{\ell}}_2 \leq \norm{\ol{x}_k - \ol{z}_{i}}_2 + \norm{\ol{z}_i - \ol{z}_j}_2 +  \norm{\ol{z}_{j} - \ol{x}_{\ell} }_2 \leq \wt{r}(1 + 2\sqrt{d}) = r/3.
-\end{equation*}
-Since we also assume $r/3 \geq d_{\infty}(X,\ol{X})$, another application of the triangle inequality gives
-\begin{equation*}
-\norm{\pi(\ol{x}_k) - \pi(\ol{x}_\ell)}_2 \leq \norm{\pi(\ol{x}_k) - \ol{x}_\ell}_2 + \norm{\ol{x}_k - \ol{x}_\ell}_2 \leq \norm{\ol{x}_{\ell} - \pi(\ol{x}_\ell)}_2 \leq r,
-\end{equation*}
-implying that $(\pi(\ol{x}_k), \pi(\ol{x}_{\ell})) \in E$ and consequently that $G_{n,r} \succeq \ol{G}^M_d \otimes K_N$.
-
-The eigenvalues of lattices and complete graphs are known to satisfy, respectively
-\begin{equation*}
-\lambda_k(\ol{G}^{M}_d) \geq \frac{k^{2/d}}{M^{2/d}}~\textrm{for $k = 0,\ldots,M - 1$},~~ \textrm{and}~\lambda_{j}(K_N) \geq N\1\{j > 0\}~\textrm{for $j = 0,\ldots,N-1$.}
-\end{equation*}
-and by standard facts regarding the eigenvalues of tensor product graphs, we have that the spectrum $\Lambda(H)$ satisfies
-\begin{equation*}
-\Lambda(H) = \set{N\lambda_k(\ol{G}^{M}_d) + M\lambda_j(K_N): \textrm{for $k = 0,\ldots,M - 1$ and $j = 0,\ldots,N-1$}}
-\end{equation*}
-For all $j = 1,\ldots,N-1$, we have that $M\lambda_j(K_N) = MN = n$. Therefore,
-\begin{align*}
-\lambda_{\kappa}(H) & \geq \{n \wedge N\lambda_{\kappa}(\ol{G}^{M}_d)\} \\
-& \geq \{n \wedge n\wt{r}^d\frac{\kappa^{2/d}}{M^{2/d}}\} \\
-& \geq \{n \wedge (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d}\} \\
-& \geq (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d},
-\end{align*}
-where the last inequality is satisfied since $r \leq \kappa^{-2/(d(d + 2))}$, completing the proof of~\eqref{eqn:eigenvalue_tail_bound}.
-
-\subsubsection{Proof of \eqref{eqn:l2_to_empirical_norm_order1}}
-The inequality~\eqref{eqn:l2_to_empirical_norm_order1} is implied by the more general Lemma~\ref{lem:empirical_norm_sobolev}, which we state in Section~\ref{subsec:sobolev_testing_rate_pf} and use to prove a similar inequality when $f \in H^s(\Xset)$ for general $s$.
-
-\subsubsection{Putting the pieces together}
-
-The final proof of Theorem~\ref{thm:sobolev_testing_rate_order1} proceeds by taking $s = 1$ and proceeding as in Section~\ref{subsubsec:sobolev_testing_rate_pf_conclusion}.
-
-\subsection{Proof of Theorem~\ref{thm:sobolev_testing_rate}}
-\label{subsec:sobolev_testing_rate_pf}
-
-We want to apply Lemma~\ref{lem:fixed_graph_testing} to the case where $G = G_{n,r}$.
-In order to do this, we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = G_{n,r}$ and $\beta = (f(x_1),\ldots,f(x_n))$.
-In particular, we will show that the following statements each hold with probability at least $1 - c/b$ for sufficiently large $n$ (Here and in what follows, $c$ denotes a constant which is fixed in $n$ and $f$ and does not depend on $b$, but may depend on other fixed quantities such as $d$, $s$, etc.):
-\begin{enumerate}[label=(E\arabic*)]
-	\item 
-	\label{event:discrete_sobolev_norm}
-	\textbf{Graph Sobolev norm:} For any $n^{-1/(2(s - 1) + d)}\leq r$ such that $r \to 0$ as $n \to \infty$,
-	\begin{equation}
-	\label{eqn:continuous_to_discrete_sobolev_norm}
-	f^T L^s f \leq c \cdot b \cdot \norm{f}_{H^s(\Xset)}^2 \cdot n^{s + 1} r^{s(d + 2)}.
-	\end{equation}
-	for all $n$ sufficiently large.
-	\item 
-	\label{event:eigenvalue_tail_decay_2}
-	\textbf{Eigenvalue tail bound:} There exists a constant $c$ such that for any $\kappa = 1,\ldots,n$, if $\max\{ 3 d_{\infty}(X,\ol{X}),n^{-1/d}\} \leq r \leq \kappa^{-2/(d(2 +d))}$ then,
-	\begin{equation}
-	\label{eqn:eigenvalue_tail_bound_2}
-	\lambda_{\kappa} \geq c \cdot n r^{d + 2} \kappa^{2/d}
-	\end{equation}
-	In particular, the inequality~\eqref{eqn:eigenvalue_tail_bound_2} is satisfied for sufficiently large $n$ when $\kappa = n^{2d/(4s + d)}$ and $c(\log n/n)^{1/d} \leq r \leq n^{-4/((2+d)(4s + d))}$. 
-	\item 
-	\label{event:l2_norm_order1}
-	\textbf{Empirical norm of $f$:} There exists a constant $c_1$ such that if $\norm{f}_{\Leb^2(\Xset)} \geq c_1 \cdot b \cdot n^{-2s/(4s + d)}$ and $4s > d$, then
-	\begin{equation}
-	\label{eqn:l2_to_empirical_norm_order1}
-	\norm{f}_n^2 \geq \frac{1}{b} \cdot \norm{f}_{\Leb^2}^2
-	\end{equation}
-\end{enumerate}
-
-\subsubsection{Proof of~\eqref{eqn:continuous_to_discrete_sobolev_norm}}
-
-The probabilistic bound~\eqref{eqn:continuous_to_discrete_sobolev_norm} follows from the more general Lemma~\ref{lem:roughness_functional_expectation_sobolev} by Markov's inequality. 
-\begin{lemma}
-	\label{lem:roughness_functional_expectation_sobolev}
-	Let $\Xset$ be a Lipschitz domain and let $s \geq 1$ be an integer.
-Suppose that $f \in H_0^{s}(\Xset)$, and further that $p \in C^{s-1}(\Xset;p_{\max})$ for some constant $p_{\max} < \infty$.
-Then for any $2$nd-order kernel $K$ and any $n^{-1/(2(s - 1) + d)} \leq r(n)$ such that $r \to 0$ as $n \to \infty$, the expected graph Sobolev seminorm is upper bounded
-	\begin{equation}
-	\label{eqn:roughness_functional_expectation_sobolev}
-	\Ebb\bigl[f^T L^s f\bigr] \leq c \cdot \norm{f}_{H^s(\Xset)}^2 \cdot n^{s + 1}r^{s(d + 2)}
-	\end{equation}
-	for all $n$ sufficiently large.
-\end{lemma}
-
-We note that the proof of Lemma~\ref{lem:roughness_functional_expectation_sobolev} is where we rely on the fact that $f$ and its derivatives are compactly supported on $\Xset$.
-The proof of Lemma~\ref{lem:roughness_functional_expectation_sobolev} is lengthy, and we defer it to Section~\ref{sec:technical_lemma_proofs}.
-
-\subsubsection{Proof of~\eqref{eqn:eigenvalue_tail_bound}}
-See Section~\ref{subsubsec:eigenvalue_tail_bound_pf}.
-
-\subsubsection{Proof of~\eqref{eqn:l2_to_empirical_norm}}
-
-\begin{lemma}
-	\label{lem:empirical_norm_sobolev}
-	Let $\Xset$ be a Lipschitz domain over which the density is upper and lower bounded 
-	\begin{equation*}
-	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty ~~\textrm{for all $x \in \Xset$,}
-	\end{equation*}
-	and let $f \in H^s(\Xset)$.Then for any $b \geq 1$, there exists $c_1$ such that if 
-	\begin{equation}
-	\norm{f}_{\Leb^2(\Xset)} \geq 
-	\begin{cases*}
-	c_1 \cdot b \cdot \norm{f}_{H^s(\Xset)} \cdot \max\{n^{-1/2},n^{-s/d}\},~~\textrm{if}~2s \neq d \\
-	c_1 \cdot b \cdot \norm{f}_{H^s(\Xset)} \cdot n^{-a/2},~~\textrm{if}~ 2s = d ~\textrm{for any}~ 0 < a < 1
-	\end{cases*}
-	\end{equation}
-	then,
-	\begin{equation}
-	\label{eqn:l2_to_empirical_norm_sobolev}
-	\mathbb{P}\left[\norm{f}_n^2 \geq \frac{1}{b} \Ebb\bigl[\norm{f}_n^2\bigr]\right] \geq 1 - \frac{5}{b}
-	\end{equation}
-	where $c_1$ and $c_2$ are constants which may depend only on $s$, $\Xset$, $d$, $p_{\min}$ and $p_{\max}$.
-\end{lemma}
-The lower bound~\eqref{eqn:l2_to_empirical_norm} results from the more general Lemma~\ref{lem:empirical_norm_sobolev}, which can be verified by checking the various orderings of $2s/(4s + d)$, $s/d$ and $1/2$ whenever $4s < d$.
-
-\paragraph{Proof of Lemma~\ref{lem:empirical_norm_sobolev}:}
-
-To prove~\eqref{eqn:l2_to_empirical_norm_sobolev} we will show
-\begin{equation*}
-\mathbb{E}\bigl[\norm{f}_n^4\bigr] \leq \left(1 + \frac{1}{b^2}\right)\cdot \left(\mathbb{E}\bigl[\norm{f}_n^2\bigr]\right)^2
-\end{equation*}
-whence the claim follows from the Paley-Zygmund inequality (Lemma~\ref{lem:paley_zygmund}).
-Since $p \leq p_{\max}$ is uniformly bounded, we can relate $\mathbb{E}\bigl[\norm{f}_n^4\bigr]$ to the $\Leb^4$ norm,
-\begin{equation*}
-\mathbb{E}\bigl[\norm{f}_n^4\bigr] = \frac{(n-1)}{n}\left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + \frac{\mathbb{E}\Bigl[\bigl(f(x_1)\bigr)^4\Bigr]}{n} \leq \left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + p_{\max}^2\frac{\norm{f}_{\Leb^4}^4}{n}.
-\end{equation*}
-We will use a Sobolev inequality to relate $\norm{f}_{\Leb^4}$ to $\norm{f}_{W_d^{s,2}(\Xset)}$.
-The nature of this inequality depends on the relationship between $s$ and $d$ (see Theorem 6 in Section 5.6.3 of \textcolor{red}{Evans} for a formal statement), so from this point on we divide our analysis into three cases: (i) the case where $2s > d$, (ii) the case where $2s < d$, and (iii) the borderline case $2s = d$.
-
-\textit{Case 1: $2s > d$.}
-When $2s > d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in C^{\gamma}(\overline{\Xset})$ for some $\gamma > 0$ which depends on $s$ and $d$, with the accompanying estimate
-\begin{equation*}
-\sup_{x \in \Xset} \abs{f(x)} \leq \norm{f}_{C^{\gamma}(\Xset)} \leq c \norm{f}_{H^s(\Xset)}.
-\end{equation*}
-Therefore,
-\begin{align*}
-\norm{f}_{\Leb^4}^4 & = \int_{\Xset} [f(x)]^4 \,dx \\
-& \leq \left(\sup_{x \in \Xset} \abs{f(x)}\right)^2 \cdot \int_{\Xset} [f(x)]^2 \,dx \\
-& \leq c \norm{f}_{H^s(\Xset)}^2 \cdot \norm{f}_{\Leb^2(\Xset)}^2.
-\end{align*}
-Since by assumption
-\begin{equation*}
-\norm{f}_{\Leb^2(\Xset)}^2 \geq c_1^2 \cdot b^2 \cdot \norm{f}_{W_d^{s,2}(\Xset)}^2 \cdot \frac{1}{n},
-\end{equation*}
-we have
-\begin{equation*}
-p_{\max}^2\frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c \norm{f}_{H^s(\Xset)}^2 \cdot \frac{\norm{f}_{\Leb^2(\Xset)}^4}{n \norm{f}_{\Leb^2(\Xset)}^2} \leq c \frac{\norm{f}_{\Leb^2(\Xset)}^4}{c_1^2 b^2} \leq \frac{\Ebb\bigl[\norm{f}_n^2\bigr]}{b^2},
-\end{equation*}
-where the last inequality follows by taking $c_1$ sufficiently large.
-
-\textit{Case 2: $2s < d$.}
-When $2s < d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in \Leb^q(\Xset)$ for $q = 2d/(d - 2s)$, and moreover that
-\begin{equation*}
-\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{H^s(\Xset)}.
-\end{equation*}
-Since $4 = 2\theta + (1 - \theta)q$ for $\theta = 2 - d/(2s)$, Lyapunov's inequality implies
-\begin{equation*}
-\norm{f}_{\Leb^4(\Xset)}^4 \leq \norm{f}_{\Leb^2}^{2\theta} \cdot \norm{f}_{\Leb^q(\Xset)}^{(1 - \theta)q} \leq c \norm{f}_{\Leb^2(\Xset)}^{4} \cdot \left(\frac{\norm{f}_{H^s(\Xset)}}{\norm{f}_{\Leb^2(\Xset)}}\right)^{d/s}.
-\end{equation*}
-By assumption, $\norm{f}_{\Leb^2(\Xset)} \geq c_1 b \norm{f}_{H^s(\Xset)} n^{-s/d}$, and therefore
-\begin{equation*}
-p_{\max}^2 \frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c\norm{f}_{\Leb^2(\Xset)}^4 \left(\frac{\norm{f}_{H^s(\Xset)}}{n^{s/d}\norm{f}_{\Leb^2(\Xset)}}\right)^{d/s} \leq \frac{c\norm{f}_{\Leb^2(\Xset)}^4}{c_1b^{d/s}} \leq \frac{\norm{f}_{\Leb^2(\Xset)}^4}{b^2}.
-\end{equation*}
-where the last inequality follows when $c_1$ is sufficiently large, and keeping in mind that $d/s > 2$ and $b \geq 1$.
-
-\textit{Case 3: $2s = d$.}
-Assume $f$ satisfies~\eqref{eqn:paley_zygmund_1} for a given $0 < a < 1$.
-When $2s = d$, since $\Xset$ is a Lipschitz domain we have that $f \in L^q(\Xset)$ for any $q < \infty$, with the accompanying estimate
-\begin{equation*}
-\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{H^s(\Xset)}.
-\end{equation*}
-In particular the above holds for $q = 2/(1 - a)$ when $1/2 < a < 1$, and for any $q > 4$ when $0 < a < 1/2$. Using Lyapunov's inequality as in the previous case then implies the desired result.
-
-\subsubsection{Putting the pieces together}
-\label{subsubsec:sobolev_testing_rate_pf_conclusion}
-
-We note that for all possible values of $X \in \Xset^n$, under the null hypothesis $f = f_0 = 0$ and therefore $\beta = (f(x_1),\ldots,f(x_n)) = 0$ as well.
-Therefore by~\eqref{eqn:graph_spectral_type_I_error}, we have the following bound on Type I error:
-\begin{equation}
-\Ebb_{f_0}(\phi_{\mathrm{spec}}) = \mathbb{E}(\mathbb{E}_{\beta = 0}(\phi_{\spec}) | X) \leq \frac{1}{b^2}.
-\end{equation}
-
-Now, we bound Type II error under the assumption $f \in H_0^{s}(\mathcal{X};R), p \in C^{s - 1}(\Xset;p_{\max})$ uniformly bounded away from $0$ and $\infty$ over $\Xset$, and 
-\begin{equation}
-\label{eqn:critical_radius_1}
-\norm{f}_{\Leb^2(\Xset)}^2 \geq \epsilon^2 = c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4s + d)}\} \cdot n^{-4s/(4s + d)}.
-\end{equation}
-Choosing $n^{-1/(2(s - 1) + d)}\leq r(n) \leq n^{-4/((2+d)(4s + d))}$, we may therefore apply our conclusions in Step 2; namely, that for every possible choice of $f \in H^s(\Xset;R)$ there exists a good set $\mathcal{E}_f \subseteq \Xset^n$ with $\Pbb(\mathcal{E}_f) \geq 1 - c/b$ such that each of \eqref{eqn:continuous_to_discrete_sobolev_norm}, \eqref{eqn:eigenvalue_tail_bound}, and \eqref{eqn:l2_to_empirical_norm_sobolev} hold for all $X \subseteq \mathcal{E}_f$.
-The choice $\kappa = (nR^2)^{2d/(4s + d)}$ balances the squared bias and variance terms on the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius}, and we have that for all $X \subseteq \mathcal{E}_f$
-\begin{align*}
-2b\sqrt{\frac{2\kappa}{n^2}} + \frac{f^T L^s f}{n\lambda_{\kappa}^s} & \leq 2bR^{2d/(4s + d)}n^{-4s/(4s+d)} + cbR^2\frac{n^{s} r^{s(d + 2)}}{\lambda_{\kappa}^s} \tag{by \eqref{eqn:continuous_to_discrete_sobolev_norm}} \\
-& \leq cbR^{2d/(4s + d)}n^{-4s/(4s+d)} \tag{by \eqref{eqn:eigenvalue_tail_bound}} \\
-& \leq \frac{1}{b}\norm{f}_{\Leb^2(\Xset)}^2 \\
-& \leq \frac{1}{n}\sum_{i = 1} \bigl[f(x_i)\bigr]^2. \tag{by \eqref{eqn:l2_to_empirical_norm_sobolev}}
-\end{align*}
-where the last two inequalities follow for a suitably large choice of $c_1$ in~\eqref{eqn:critical_radius_1}.
-We conclude that for all $X \subseteq \mathcal{E}_f$, the inequality \eqref{eqn:fixed_graph_testing_critical_radius} is satisfied with respect to $\beta = (f(x_1),\ldots,f(x_n))$ and $G = G_{n,r}$.
-As a result the worst-case Type II error is bounded
-\begin{equation*}
-\sup_{\substack{f \in H^s(\Xset;R), \\ \norm{f}_{\Leb^2(\Xset)} \geq \epsilon}}\mathbb{E}_{f}(1 - \phi_{\spec}) \leq \sup_{\substack{f \in H^s(\Xset;R), \\ \norm{f}_{\Leb^2(\Xset)} \geq \epsilon}} \mathbb{E}\bigl[\mathbb{E}_{\beta}(1 - \phi_{\spec}|X \in \mathcal{E}_f)\bigr] + \frac{c}{b} \leq \frac{3 + c}{b},
-\end{equation*}
-completing the proof of Theorem~\ref{thm:sobolev_testing_rate} upon proper choice of constants $c_1,c_2$ in that Theorem.
-
-\subsection{Proof of Proposition~\ref{prop:L4_testing_rate}}
-
-Let $T_{\mathrm{mean}} = \frac{1}{n}\sum_{i = 1}^{n} y_i^2$. The expectation of $T_{\mathrm{mean}}$ is
-\begin{equation*}
-\Ebb(T_{\mathrm{mean}}) = \mathbb{E}(f^2(x_1)) + 1,
-\end{equation*}
-and the variance can be upper bounded
-\begin{equation*}
-\Var(T_{\mathrm{mean}}) \leq \frac{1}{n}(3 + p_{\max} \norm{f}_{\Leb^4}^4 + p_{\max}\norm{f}_{\Leb^2}^2) = \frac{c}{n}.
-\end{equation*}
-Since $\mathbb{E}(f^2(x_1)) \geq p_{\min} \epsilon^2$, when $\epsilon^2 \gtrsim b\cdot n^{-1/2}$ we can apply Chebyshev's inequality to obtain the claimed result.
-
-\subsection{Proof of Theorem~\ref{thm:twosample_sobolev_testing_rate}}
-
-As mentioned previously, to prove Theorem~\ref{thm:twosample_sobolev_testing_rate}, we relate the two sample model to the following regression model: we observe
-$X = \{x_1,\ldots,x_n\} \sim \mu$ (where we recall $\mu = (P + Q)/2$), and associated labels
-\begin{equation*}
-a_i = 
-\begin{cases}
-1, & \textrm{with probability}~ \frac{p(x_i)}{p(x_i) + q(x_i)} \\
--1, & \textrm{with probability}~ \frac{q(x_i)}{p(x_i) + q(x_i)}
-\end{cases}
-\end{equation*}
-where $a_i$ is conditionally independent of $x_j,a_j$ given $x_i$.
-
-\textcolor{red}{TODO:} Complete the above argument.
-
-Our analysis will trace a similar path to the proof of Theorem~\ref{thm:sobolev_testing_rate}, proceeding according to the following steps:
-\begin{enumerate}
-	\item We upper bound the testing error of our eigenvector projection test statistic when $a$ are viewed as random responses defined over the vertices of a fixed graph $G$.
-This upper bound will hold whenever certain functionals on the graph $G$ are themselves bounded.
-One of these functionals will be a measure of eigenvector incoherence.
-	\item We analyze the behavior of these functionals with respect to the random graph $G_{n,r}$, and bound them with high probability.
-	\item We condition on $X \subseteq \mathcal{E}$, where $\mathcal{E}$ is a high-probability set over which the relevant functionals on $G_{n,r}$ are bounded.
-We conclude that the upper bound on testing error derived in our first step holds whenever $X \subseteq \mathcal{E}$.
-\end{enumerate}
-
-\subsubsection{Step 1: Testing error on a fixed graph}
-
-Let $G = (V,E)$ be a graph over vertices $V = \set{v_1,\ldots,v_n}$, and let $\beta^{(p)} = (\beta_1^{(p)},\ldots,\beta_n^{(p)}) \in \Reals^n$, $\beta^{(q)} = (\beta_1^{(q)},\ldots,\beta_n^{(q)}) \in \Reals^n$ be non-negative signals over the vertices $V$.
-We observe labels $a = (a_1,\ldots,a_n)$ according to the model
-\begin{equation*}
-a_i = 
-\begin{cases*}
-1, & \textrm{with probability $\frac{\betap_i}{\betap_i + \betaq_i}$} \\
--1, & \textrm{with probability $\frac{\betaq_i}{\betap_i + \betaq_i}$}
-\end{cases*}
-\end{equation*}
-Letting $L = VSV^T$ be the spectral decomposition of the Laplacian $L$ of $G$, our graph spectral test statistic is 
-\begin{equation*}
-T_{\spec}^{(2)} = \frac{1}{n}\sum_{k = 1}^{\kappa} \left(\sum_{i = 1}^{n} v_i a_i\right)^2
-\end{equation*}
-where $\kappa$ is a tuning parameter. The resulting test we will use is
-\begin{equation*}
-\phi_{\spec}^{(2)} = \1\{T_{\spec}^{(2)} \geq \frac{\kappa}{n} + t(b)\},~~\textrm{where}~ t(b) = b\sqrt{\frac{2\kappa}{n^2}}~~\textrm{for $b \geq 1$.}
-\end{equation*}
-
-Let $\Pi_{\max}(\kappa;G)$ be a measure of the incoherence of the eigenvectors $v_1,\ldots,v_\kappa$, given by
-\begin{equation*}
-\Pi_{\max}(\kappa;G) := \max_{i = 1,\ldots,n} \left\{\sum_{k = 1}^{\kappa} v_{k,i}^2\right\}
-\end{equation*}
-In Lemma~\ref{lem:twosample_fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}^{(2)}$.
-Our bound on the Type II error will be stated as a function of $\Pi_{\max}(\kappa;G)$ as well as the $\kappa$th eigenvalue $s_{\kappa}$ and the smoothness functional $S_2(\beta;G)$.
-We use the notation $\varDelta^{(p,q)} = (\betap - \betaq)/(\betap + \betaq)$.
-
-\begin{lemma}
-	\label{lem:twosample_fixed_graph_testing}
-	Let $1 \leq \kappa \leq n$ be an integer.
-	\begin{enumerate}
-		\item \textbf{Type I error:} Under the null hypothesis $\betap = \betaq$, the Type I error of $\phi_{\spec}^{(2)}$ is upper bounded
-		\begin{equation}
-		\label{eqn:twosample_graph_spectral_type_I_error}
-		\mathbb{E}_{\betap,\betap}(\phi_{\spec}^{(2)}) \leq \frac{1}{b^2}.
-		\end{equation}
-		\item \textbf{Type II error:} Suppose $\Pi_{\max}(\kappa;G) \leq 1$. Then for any $b$, $\betap$ and $\betaq$ such that
-		\begin{equation}
-		\label{eqn:twosample_fixed_graph_testing_critical_radius}
-		\frac{1}{n} \sum_{i = 1}^{n} (\varDelta_i^{(p,q)})^2 \geq \frac{1}{1 - \Pi_{\max}(\kappa;G)}\left(2b\sqrt{\frac{2\kappa}{n^2}} + \frac{S_2(\varDelta;G)}{ns_{\kappa}}\right)
-		\end{equation}
-		the Type II error of $\phi_{\spec}^{(2)}$ is upper bounded,
-		\begin{equation}
-		\label{eqn:twosample_graph_spectral_type_II_error}
-		\mathbb{E}_{\betap,\betaq}(1 - \phi_{\spec}^{(2)}) \leq \frac{3}{b}.
-		\end{equation}
-	\end{enumerate}
-\end{lemma}
-
-\paragraph{Proof of Lemma~\ref{lem:twosample_fixed_graph_testing}.}
-
-To prove Lemma~\ref{lem:twosample_fixed_graph_testing} we will compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}^{(2)}$.
-The inequalities \eqref{eqn:twosample_graph_spectral_type_I_error} and \eqref{eqn:twosample_graph_spectral_type_II_error} can then be computed using Chebyshev's inequality in a manner very similar to the proof of Lemma~\ref{lem:fixed_graph_testing}, and we omit the details.
-Let $w_i := a_i - \vardeltapq_{i}$, and note that $\Ebb(w_i) = 0$ and
-\begin{equation*}
-\Ebb(w_i^2) = \frac{\betap_i \betaq_i}{(\betap_i + \betaq_i)^2} = 1 - (\vardeltapq_{i})^2,~~ \Ebb(w_i^3) = \frac{1}{2}(1 - (\vardeltapq_{i})^2)\vardeltapq_{i}
-\end{equation*}
-
-\vspace{.2 in}
-
-\textit{Mean of $T_{\spec}^{(2)}$}:
-
-We have
-\begin{align*}
-\Ebb(T_{\spec}^{(2)}) & = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{\vardeltapq}{v_k}^2 + \Ebb\bigl(\dotp{w}{v_k}^2 + 2 \dotp{w}{v_k} \dotp{\vardeltapq}{v_k}\bigr)\right) \\
-& = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{1 - (\vardeltapq)^2}{v_k^2} + \dotp{\vardeltapq}{v_k}\right) \\
-& \geq \frac{\kappa}{n} + \left(\sum_{k = 1}^{\kappa}\dotp{\vardeltapq}{v_k} - \dotp{(\vardeltapq)^2}{v_k^2}\right)
-\end{align*}
-When $\vardeltapq = 0$, this equals $\kappa/n$. Otherwise, we have the following upper bound:
-\begin{align*}
-\sum_{k = 1}^{\kappa} \dotp{(\vardeltapq)^2}{v_k^2}  & = \sum_{i = 1}^{n} (\vardeltapq_{i})^2 \left(\sum_{k = 1}^{\kappa} v_{k,i}^2 \right)\\
-& \leq \Pi_{\max}(\kappa,G) \cdot \sum_{i = 1}^{n} (\vardeltapq_{i})^2 = \Pi_{\max}(\kappa,G) \cdot \norm{\vardeltapq}_2^2.
-\end{align*}
-Additionally, $\sum_{k = 1}^{\kappa}\dotp{\vardeltapq}{v_k} \geq \norm{\vardeltapq}_2^2 - S_2(\beta;G)/s_{\kappa}$ by reasoning given in the proof of Lemma~\ref{lem:fixed_graph_testing}, and as a result
-\begin{equation*}
-\Ebb(T_{\spec}^{(2)}) \geq \frac{\kappa}{n} -  \frac{\bigl(1 - \Pi_{\max}(\kappa;G)\bigr)}{n} \norm{\vardeltapq}_2^2 - \frac{S_2(\beta;G)}{ns_{\kappa}}
-\end{equation*}
-
-\vspace{.2 in}
-\textit{Variance of $T_{\spec}^{(2)}$:}
-
-We have
-\begin{align}
-\Var(T_{\spec}^{(2)}) & = \frac{1}{n^2}\Var(2\dotp{V_{\kappa}V_{\kappa}^T\vardeltapq}{w} + \dotp{V_{\kappa}V_{\kappa}^Tw}{w}) \nonumber \\
-& = \frac{1}{n^2}\Bigl\{4\underbrace{\Var(\dotp{V_{\kappa}V_{\kappa}^T\vardeltapq}{w})}_{=: V_1} + 2\underbrace{\Cov(\dotp{V_{\kappa}V_{\kappa}^T\vardeltapq}{w},\dotp{V_{\kappa}V_{\kappa}^Tw}{w})}_{=: K_1} + \underbrace{\Var(\dotp{V_{\kappa}V_{\kappa}^Tw}{w})}_{=:V_2}\Bigr\}, \label{eqn:var}
-\end{align}
-and we now upper bound each of the three terms on the right hand side of the previous display.
-
-\textbf{Upper bound on $V_1$:}
-Let $\Sigma := \Cov(w)$ be the covariance matrix of $w$. Noting that $\Sigma \preceq I$, we have
-\begin{equation}
-\label{eqn:var1}
-\Var(\dotp{V_{\kappa}V_{\kappa}^T\vardeltapq}{w}) =(\vardeltapq)^T V_{\kappa}V_{\kappa}^T \Sigma V_{\kappa}V_{\kappa}^T \vardeltapq \leq \norm{V_{\kappa}V_{\kappa}^T\vardeltapq}^2.
-\end{equation}
-
-\textbf{Upper bound on $K_1$:}
-Noting that $\Ebb(\dotp{V_{\kappa}V_{\kappa}^T \vardeltapq}{w}) = 0$, we have that
-\begin{align}
-K_1 & = \Ebb\left[\dotp{V_{\kappa}V_{\kappa}^T \vardeltapq}{w} \dotp{V_{\kappa}V_{\kappa}^T w}{w}\right] \nonumber \\
-& = \sum_{i = 1}^{n} \sum_{j = 1}^{n} \sum_{i' = 1}^{n} \Ebb\left[w_i w_j (V_{\kappa}V_{\kappa}^T)_{ij} w_{i'} (V_{\kappa}V_{\kappa}^T\vardeltapq)_i\right] \nonumber \\
-& = \sum_{i = 1}^{n} \Ebb\left[w_i^3\right] (V_{\kappa}V_{\kappa}^T)_{ii} (V_{\kappa}V_{\kappa}^T \vardeltapq)_i \nonumber \\
-& = \frac{1}{2}\sum_{i = 1}^{n} (1 - (\vardeltapq_{i})^2)(\vardeltapq_{i}) (V_{\kappa} V_{\kappa}^T)_{ii}  (V_{\kappa}V_{\kappa}^T\vardeltapq)_i \nonumber \\
-& \leq \norm{V_{\kappa}V_{\kappa}^T \vardeltapq} \cdot \left(\sum_{i = 1}^{n} (\vardeltapq_{i})^2 (V_{\kappa} V_{\kappa}^T)_{ii}^2\right)^{1/2} \nonumber \\
-& \leq \Pi(\kappa,G) \cdot \norm{V_{\kappa}V_{\kappa}^T \vardeltapq} \cdot \norm{\vardeltapq}  \label{eqn:var2} 
-\end{align}
-
-\textbf{Upper bound on $V_2$:}
-$V_2$ is a variance of a sum, which we re-express as the sum of covariances:
-\begin{align*}
-V_2 & = \Var(\dotp{V_{\kappa}V_{\kappa}^T w}{w}) \\
-& = \sum_{i,j,i',j' = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ij} (V_{\kappa}V_{\kappa}^T)_{i'j'} \Cov(w_i w_j, w_{i'} w_{j'}).
-\end{align*}
-This covariance will be non-zero only when $i = i' \neq j = j'$, $i = j' \neq j = i'$, or $i = i' = j = j'$, and therefore
-\begin{align}
-V_2 & = 2\sum_{i = 1}^{n}\sum_{j = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ij}^2 \Var(w_iw_j) + \sum_{i = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ii}^2 \Var(w_i^2) \nonumber\\
-& \leq 2\sum_{i = 1}^{n}\sum_{j = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ij}^2 + \sum_{i = 1}^{n} (V_{\kappa}V_{\kappa}^T)_{ii}^2 \leq 3 \mathrm{tr}\bigl((V_{\kappa}V_{\kappa}^T)^2\bigr) = 3\kappa.
-\label{eqn:var3}
-\end{align}
-
-Now, plugging \eqref{eqn:var1}, \eqref{eqn:var2}, and \eqref{eqn:var3} back into \eqref{eqn:var}, we obtain
-\begin{equation}
-\label{eqn:var_5}
-\Var(T_{\spec}^{(2)}) \leq \frac{1}{n^2}\left\{4\norm{V_{\kappa}V_{\kappa}^T \vardeltapq}^2 + \Pi(\kappa;G) \cdot \norm{V_{\kappa}V_{\kappa}^T \vardeltapq} \cdot \norm{\vardeltapq} + 3\kappa \right\}.
-\end{equation}
-
-\subsubsection{Step 2: Bounding neighborhood graph functionals}
-
-Note that since $p,q \in H^1(\mathcal{X};R)$, we have $f := p - q \in H^1(\mathcal{X};2R)$.
-Therefore the bounds~\eqref{eqn:continuous_to_discrete_sobolev_norm} and \eqref{eqn:l2_to_empirical_norm} apply to $S_2(\varDelta,G_{n,r})$ and $(1/n)\sum_{i = 1}^{n}\varDelta_i^2$, respectively.
-The bound~\eqref{eqn:eigenvalue_tail_bound} continues to apply to $s_{\kappa}$.
-What remains is to upper bound the incoherence parameter
-\begin{equation}
-\label{eqn:pi_max}
-\Pi_{\max}(\kappa,G_{n,r}) \leq \frac{1}{2}~~\textrm{when}~ \kappa = n^{2/5}~\textrm{and}~ r = \log^a n \cdot \left(\frac{\log n}{n}\right)
-\end{equation}
-with probability $1 - o(1)$ as $n \to \infty$.
-
-Our proof of~\eqref{eqn:pi_max} will proceed by relating $G_{n,r}$ to a graph which satsifies a strict notion of incoherence, the chain graph.
-When $G = \overline{G}$, the incoherence parameter $\Pi_{\max}$ is upper bounded
-\begin{equation}
-\Pi_{\max}(\kappa,\overline{G}) \leq \frac{\kappa}{n}.
-\end{equation}
-To prove that the weaker bound \eqref{eqn:pi_max} holds with respect to $G_{n,r}$, we will show that the two graphs $\overline{G}$ and $G_{n,r}$ are $\delta$-spectrally similar, meaning
-\begin{equation}
-\label{eqn:spectral_similarity}
-(1 - \delta_n) x^T L x \leq x^T \overline{L} x \leq (1 + \delta_n) x^T L x~~\textrm{for all}~x \in \Reals^n, \delta_n = c \cdot (\log n)^{h(a,1) + da}\cdot \frac{\log n}{n}
-\end{equation}
-with probability $1 - o(1)$ as $n \to \infty$.
-The proof of~\eqref{eqn:spectral_similarity} relies heavily on bounding the transport distance between $X$ and $\overline{X}$, i.e.
-\eqref{eqn:slepcev_transport_distance}.
-Then the following technical Lemma along with \eqref{eqn:spectral_similarity} will imply \eqref{eqn:pi_max}.
-\begin{lemma}
-	\label{lem:pi_max_pf_1}
-	Let $G$ satisfy~\eqref{eqn:spectral_similarity} for a given $0 \leq \delta < 1$. Then for any $R \gtrsim n \delta^{1/2} (s_{n}^{1/2} \vee 1)$ and $k \leq n/8$,
-	\begin{equation*}
-	v_{k,i}^2 \lesssim \frac{1}{n}\left(R + \frac{n^4 \delta^2 s_{n}^2}{R^3}\right) \quad \textrm{for every $i = 1,\ldots,n$.}
-	\end{equation*} 
-\end{lemma}
-
-\textcolor{red}{TODO}: Explain how Lemma~\ref{lem:pi_max_pf_1} implies~\eqref{eqn:pi_max}.
-
-At a high level, Lemma~\ref{lem:pi_max_pf_1} is proved through repeated applications of the Davis-Kahan Theorem.
-The full proof is long and we delay presenting it until Section~\ref{sec:technical_lemma_proofs}.
-
-\subsubsection{Step 3: Conclusion}
-
-With Lemma~\ref{lem:twosample_fixed_graph_testing} as well as \eqref{eqn:continuous_to_discrete_sobolev_norm}, \eqref{eqn:eigenvalue_tail_bound}, \eqref{eqn:l2_to_empirical_norm} and \eqref{eqn:pi_max} in hand, the proof of Theorem~\ref{thm:twosample_sobolev_testing_rate} follows by similar reasoning to the conclusion of Theorem~\ref{thm:sobolev_testing_rate}.
-
-\section{Proofs of Technical Results}
-\label{sec:technical_lemma_proofs}
-
-Here we give in full detail the proofs of Lemma~\ref{lem:roughness_functional_expectation_sobolev}, \eqref{eqn:spectral_similarity}, and Lemma~\ref{lem:pi_max_pf_1}.
-
-\subsection{Proof of Lemma~\ref{lem:roughness_functional_expectation_sobolev}}
-To simplify exposition, we introduce the iterated difference operator, defined recursively as
-\begin{equation*}
-D_{jk}f(x) = (D_{k}f(x_j) - D_{k}f(x))\frac{K_r(x_j,x)}{r^d},~~ D_jf(x) = (f(x_j) - f(x))\frac{K_r(x_j,x)}{r^d}~~ \textrm{for $j \in [n], k \in [n]^q$}
-\end{equation*}
-We will also use the notation $d_jf(x) := (f(x_j) - f(x))$. We split our analysis into cases based on whether $s$ is even or odd.
-
-\subsubsection{Case 1: $s$ is even.}
-When $s$ is even, letting $q = s/2$ we have the decomposition
-\begin{equation}
-\label{eqn:continuous_to_discrete_sobolev_norm_pf1}
-f^T L^s f =  r^{ds} \cdot \sum_{i = 1}^{n} \sum_{k \in [n]^q} \sum_{\ell \in [n]^q} D_kf(x_i) D_{\ell}f(x_i). 
-\end{equation}
-
-For given index vectors $k,\ell \in [n]^q$ and indices $i,j$, let $I = \abs{k \cup \ell \cup i}$ be the total number of unique indices.
-We separate our analysis into cases based on the magnitude of $I$, specifically whether $I = s + 1$ (the leading terms where all indices are distinct) $I < s + 1$ (the terms where at least one index is repeated) and show that
-\begin{equation}
-\label{eqn:expected_difference_operators_1}
-\Ebb(D_kf(x_i) D_\ell f(x_i)) =
-\begin{cases*}
-O(r^{2s}) \cdot \norm{f}_{H^s(\Xset)}^2, & ~~\textrm{if $I = s + 1$} \\
-O(r^{2} r^{d(I - (s + 1))}) \cdot [f]_{H^1(\Xset)}^2, & ~~\textrm{if $I < s + 1$}~ 
-\end{cases*}
-\end{equation}
-We will prove~\eqref{eqn:expected_difference_operators_1} in Section~\ref{subsec:expected_difference_operators_pf}.
-First, we verify that~\eqref{eqn:continuous_to_discrete_sobolev_norm_pf1} and \eqref{eqn:expected_difference_operators_1} are together enough to show Lemma~\ref{lem:roughness_functional_expectation_sobolev} when $s$ is even.
-In the sum on the right hand side of~\eqref{eqn:continuous_to_discrete_sobolev_norm_pf1}, there are $O(n^{I})$ terms with exactly $I$ distinct indices.
-When $I < s + 1$, by~\eqref{eqn:expected_difference_operators_1} the total contribution of such terms to the sum is $O(n^{I}r^{d(I - 1) + 2}) \cdot [f]_{H^1(\Xset)}^2$.
-Since by assumption $r \geq n^{-1/d}$, this increases with $I$.
-Taking $I = s$ to be the largest integer less than $s + 1$, the contribution of these terms to the sum is therefore $O(n^sr^{d(s - 1) + 2}) \cdot [f]_{H^1(\Xset)}^2$ which in light of the restriction $r \geq n^{-1/(2(s - 1) + d)}$ is $O(n^{s+1}r^{s(d +2)}) \cdot [f]_{H^1(\Xset)}^2$.
-On the other hand when $I = s + 1$, by~\eqref{eqn:expected_difference_operators_1} we immediately have that the total contribution of these terms is $O(n^{s + 1}r^{2(s + d)}) \cdot \norm{f}_{H^s(\Xset)}$.
-Therefore,
-\begin{equation*}
-\Ebb(f^T L^s f) = O(n^{s+1}r^{s(d+2)}) \cdot \norm{f}_{H^s(\Xset)}^2.
-\end{equation*}
-
-\subsubsection{Case 2: $s$ is odd.}
-When $s$ is odd, letting $q = (s - 1)/2$ we have
-\begin{equation}
-\label{eqn:roughness_functional_representation_odd}
-f^T L^s f =  r^{ds} \cdot \sum_{i,j = 1}^{n} \sum_{k \in [n]^q} \sum_{\ell \in [n]^q} \bigl(d_jD_kf(x_i)\bigr) \cdot  \bigl(d_jD_{\ell}f(x_i)\bigr) \cdot K_r(x_i,x_j).
-\end{equation}
-For given index vectors $k,\ell \in [n]^q$ and indices $i,j \in [n]$, let $I = \abs{k \cup \ell \cup i \cup j}$ be the total number of unique indices.
-Similar to the case when $s$ is even, we show that
-\begin{equation}
-\label{eqn:expected_difference_operators_2}
-\Ebb(d_iD_kf(x_j) d_iD_\ell f(x_j)) =
-\begin{cases*}
-O(r^{2s}) \cdot \norm{f}_{H^s(\Xset)}^2, & ~~\textrm{if $I = s + 1$} \\
-O(r^{2} \cdot r^{d(I - (s + 1))}) \cdot [f]_{H^1(\Xset)}^2, & ~~\textrm{if $I < s + 1$}~ 
-\end{cases*}
-\end{equation}
-Then Lemma~\ref{lem:roughness_functional_expectation_sobolev} follows from similar reasoning to the case where $s$ was even.
-
-\subsubsection{Proof of~\eqref{eqn:expected_difference_operators_1} and~\eqref{eqn:expected_difference_operators_2}}
-\label{subsec:expected_difference_operators_pf}
-
-We first prove the desired bounds in the case when some indices are repeated, and then the desired bounds in the case when all indices are distinct.
-
-\paragraph{Repeated indices.}
-
-Since the proofs of~\eqref{eqn:expected_difference_operators_1} and~\eqref{eqn:expected_difference_operators_2} are essentially the same for the case where some index is repeated, we will assume without loss of generality that $s$ is even.
-Let $k,\ell \in [n]^q$ be index vectors for $q = s/2$.
-
-When at least one index is repeated, we obtain a sufficient upper bound by reducing the problem of upper bounding the iterated difference operator to that of upper bounding a single difference operator.
-Letting $k = (k_1,\ldots,k_q)$, we can show by induction that the absolute value of the iterated difference operator $\abs{D_kf(x_i)}$ is upper bounded by
-\begin{equation*}
-\abs{D_kf(x_i)} \leq \left(\frac{2K_{\max}}{r^d}\right)^{q-1} \sum_{h \in k \cup i} \abs{D_{k_q}f(x_h)} \cdot \1\{G_{n,r}[X_{k \cup i}]~\textrm{is a connected graph} \}.
-\end{equation*}
-Therefore,
-\begin{align}
-\abs{D_kf(x_i)} \cdot \abs{D_{\ell}f(x_i)} & \leq \left(\frac{2K_{\max}}{r^d}\right)^{2(q - 1)} \sum_{h,j \in k \cup \ell \cup i} \abs{D_{k_q}f(x_h)} \cdot \abs{D_{\ell_q}f(x_j)} \cdot \1\{G_{n,r}[X_{k \cup i}], G_{n,r}[X_{\ell \cup i}]~\textrm{are connected graphs.} \} \nonumber \\
-& =  \left(\frac{2K_{\max}}{r^d}\right)^{2(q - 1)} \sum_{h,j \in k \cup \ell \cup i} \abs{D_{k_q}f(x_h)} \cdot \abs{D_{\ell_q}f(x_j)} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is a connected graph.} \} \label{eqn:expected_difference_operators_sobolev_pf0}
-\end{align}
-
-We now break our analysis into three cases, based on the number of distinct indices in $k_q,\ell_q,h,j$. In each case we will obtain the same rate
-\begin{equation*}
-\Ebb\Bigl[\abs{D_{k_q}f(x_h)} \cdot \abs{D_{\ell_q}f(x_j)}\Bigr] = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot [f]_{H^1(\Xset)}^2,
-\end{equation*}
-and plugging this back in to~\eqref{eqn:expected_difference_operators_sobolev_pf0} we have that for any $k, \ell \in [n]^q$
-\begin{equation*}
-\Ebb\Bigl[\bigl|D_{k}f(x_i)\bigr| \cdot \bigl|D_{\ell}f(x_i)\bigr|\Bigr] = O(r^{(\abs{k \cup \ell \cup i} - (2q + 1))d + 2}) \cdot [f]_{H^1(\Rd)}^2.
-\end{equation*}
-
-\textit{Case 1: Two distinct indices.}
-Let $k_q = \ell_q = i$, and $h = j$. Using the law of iterated expectation, we obtain
-\begin{align*}
-\Ebb \left[\bigl(D_{i}f(x_j)\bigr)^2 \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\right] &= \Ebb \left[\bigl(D_{i}f(x_j)\bigr)^2 \cdot \Pbb\bigl[\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \} |x_i,x_j\bigr]\right] \\
-& = O(r^{(\abs{k \cup \ell \cup i} - 2)d}) \cdot \Ebb\left[\bigl(D_{i}f(x_j)\bigr)^2\right] \\
-& = O(r^{(\abs{k \cup \ell \cup i} - 3)d}) \cdot \Ebb\left[\bigl(d_{i}f(x_j)\bigr)^2K_r(x_i,x_j)\right] \\
-& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot [f]_{H^1(\Xset)}^2
-\end{align*}
-where the last equality follows from Lemma~\ref{lem:expected_first_order_seminorm}.
-
-\textit{Case 2: Three distinct indices.}
-Let $k_q = \ell_q = i$, for some $i \neq j \neq h$. Using the law of iterated expectation, we obtain
-\begin{align*}
-\Ebb \Bigl[ & \abs{D_{i}f(x_j)} \cdot \abs{D_if(x_h)} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\Bigr] = \nonumber \\
-& \Ebb\Bigl[\abs{D_{i}f(x_j)} \cdot \abs{D_if(x_h)} \cdot \Pbb\bigl[\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \} |x_i,x_j,x_h\bigr]\Bigr] \nonumber \\
-& = O(r^{(\abs{k \cup \ell \cup i} - 3)d}) \cdot \Ebb\Bigl[\abs{D_{i}f(x_j)} \cdot \abs{D_if(x_h)} \Bigr] \nonumber \\
-& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2})\cdot[f]_{H^1(\Xset)}^2
-\end{align*}
-where the last equality follows from Lemma~\ref{lem:expected_first_order_seminorm_2}.
-
-\textit{Case 3: Four distinct indices.}
-Using the law of iterated expectation, we find that
-\begin{align*}
-\Ebb\Bigl[ &\abs{D_{k_q}f(x_i)}\cdot{\abs{D_{\ell_q}f(x_j)}} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\Bigr] \\
-& = \Ebb\Bigl[\abs{D_{k_q}f(x_i)}\cdot{\abs{D_{\ell_q}f(x_j)}} \cdot\Pbb\bigl[G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected}|x_i,x_j,x_{k_q},x_{\ell_q}\bigr]\Bigr] \\
-& = O(r^{(\abs{k \cup \ell \cup i} - 4)d}) \cdot \Ebb\Bigl[\abs{D_{k_q}f(x_i)}\cdot{\abs{D_{\ell_q}f(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr] \\
-& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot [f]_{H^1(\Xset)}^2
-\end{align*}
-where the last equality follows from Lemma~\ref{lem:expected_first_order_seminorm_3}.
-
-\paragraph{All indices distinct.}
-
-We first show the desired result when $s$ is even, and then when $s$ is odd.
-
-\textit{Case 1: $s$ is even.}
-
-The result follows from the law of iterated expectation, the condition that the density be upper bounded, and equations~\eqref{eqn:leading_term_sobolev_compact_1} and \eqref{eqn:leading_term_sobolev_compact_2} in Lemma~\ref{lem:leading_term_sobolev_compact}
-\begin{equation*}
-\Ebb\bigl[D_kf(x_i)D_kf(x_j)\bigr] = \Ebb\Bigl[\bigl(\Ebb[D_kf(x_i)|x_i]\bigr)^2\Bigr] \leq p_{\max} \norm{\Ebb\bigl[D_kf\bigr]}_{\Leb^2(\Xset)}^2 \leq c r^{2s} \norm{f}_{H^s(\Xset)}^2.
-\end{equation*}
-
-\textit{Case 2: $s$ is odd.}
-
-By Lemma~\ref{lem:leading_term_sobolev_compact}, we have that for $k \in (n)^{(s - 1)/2}$, there exists a function $f_{s - 1,q} \in H_0^1(\Xset)$ such that
-\begin{equation*}
-\Bigl\|\Ebb\bigl[D_kf\bigr] - r^{2{s - 1}}f_{s - 1,q}\Bigr\|_{\Leb^2(\Xset)} \leq c r^{2s} \norm{f}_{H^s(\Rd)}^2
-\end{equation*}
-and $\norm{f_{s - 1,q}}_{H^1(\Xset)} \leq c \norm{f}_{H^s(\Xset)}$. By the law of iterated expectation, and writing $\varDelta := \Ebb\bigl[D_kf\bigr] - r^{2{s - 1}}f_{s - 1,q}$, we have
-\begin{align*}
-\Ebb\biggl[\Bigl(d_iD_kf(x_j)\Bigr) \Bigl(d_iD_{\ell}f(x_j)\Bigr) K_r(x_i,x_j)\biggr] & =  \Ebb\biggl[\Bigl(d_i\bigl(\Ebb[D_kf|x_i,x_j]\bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \nonumber \\
-& = \Ebb\biggl[\Bigl(d_i\bigl(r^{s - 1}\cdot f_{s - 1,q} + \varDelta \bigr)(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \\
-& \leq 2r^{2(s - 1)}\Ebb\biggl[\Bigl(d_i\bigl(f_{s - 1,q}\bigr)\Bigr)^2K_r(x_i,x_j)\biggr]  + 2 \Ebb \biggl[\Bigl(d_i\varDelta(x_j)\Bigr)^2K_r(x_i,x_j)\biggr] \\
-& \leq 2r^{2(s - 1)}\Ebb\biggl[\Bigl(d_i\bigl(f_{s - 1,q}\bigr)\Bigr)^2K_r(x_i,x_j)\biggr] + 2p_{\max}^2\norm{\varDelta}_{\Leb^2(\Xset)}^2 \\
-& \leq c r^{2s} \norm{f}_{H^s(\Xset)}^2
-\label{eqn:expected_difference_operators_sobolev_pf1}
-\end{align*}
-where the last inequality follows by Lemma~\ref{lem:expected_first_order_seminorm}.
-
-\subsection{Proof of~\eqref{eqn:spectral_similarity}}
-
-Assuming~\eqref{eqn:slepcev_transport_distance} holds and $r$ is chosen according to~\eqref{eqn:pi_max}, we have already shown that the upper bound in~\eqref{eqn:spectral_similarity} holds for $\delta = 0$ (see the proof of ~\eqref{eqn:eigenvalue_tail_bound}).
-Now we need only to show the lower bound, which we will do following these steps:
-\begin{enumerate}[(i)]
-	\item First, we consider the case of two fixed graphs $G$ and $\wt{G}$, and establish a Poincare inequality.
-In other words, we show that if edges in $G$ can be mapped to short, non-overlapping paths in $\wt{G}$, then the Laplacian matrix $L$ cannot be much larger $\wt{L}$, in the usual sense of partial ordering of matrices.
-	\item Then, we treat the case of $G = G_{n,r}$ and $\wt{G} = \overline{G}$.
-We assume that~\eqref{eqn:slepcev_transport_distance} holds, and exhibit such a mapping between edges in $G_{n,r}$ and paths in $\overline{G}$.
-The resulting Poincare inequality will imply~\eqref{eqn:spectral_similarity}.
-\end{enumerate}
-In step (ii) we assume that~\eqref{eqn:slepcev_transport_distance} holds.
-Since \eqref{eqn:slepcev_transport_distance} holds with probability at least $1 - c/n$, this will imply that~\eqref{eqn:spectral_similarity} holds with at least the same probability.
-
-\subsubsection{Step 1: Poincare inequality}
-
-Let $G = (V,E)$ and $\wt{G} = (V,\wt{E})$ be two graphs over a common vertex set $V$. The set $\mathcal{P}_{\wt{G}}$ consists of all paths $P$ over $\wt{G}$, meaning all tuples
-\begin{equation*}
-P = (\wt{e}_1,\wt{e}_2,\ldots,\wt{e}_m),~~m \in \mathbb{N}~, \wt{e}_j \in \wt{E}~ \forall~ j \in [m], (\wt{e}_j)_2 = (\wt{e}_{j + 1})_1 \forall~j \in [m - 1].
-\end{equation*}
-Let $\gamma:E \to \mathcal{P}_{\wt{G}}$ be a mapping from edges in $G$ to paths in $\wt{G}$. The maximum path length $M(\gamma)$ is defined
-\begin{equation*}
-M(\gamma) = \max_{e \in E} \abs{\gamma(e)}
-\end{equation*} 
-and the bottleneck $b(\gamma)$ is defined
-\begin{equation*}
-b(\gamma) = \max_{\wt{e} \in \wt{E}} \abs{\{e \in E:  \wt{e} \in \gamma(e)\}}.
-\end{equation*}
-As mentioned previously, the Laplacian matrix $L$ cannot be much larger than $\wt{L}$ if edges in $G$ can be mapped to short, non-overlapping paths in $\wt{G}$.
-Formally, we will show the following: for any $\gamma: E \to \mathcal{P}_{\wt{E}}$,
-\begin{equation}
-\label{eqn:poincare_inequality}
-x^T L x \leq M(\gamma)\cdot b(\gamma) \cdot (x^T \wt{L} x)~\forall{x \in \Reals^n}
-\end{equation}
-
-\paragraph{Proof of~\eqref{eqn:poincare_inequality}:}
-
-Letting $c \in \Reals$, we will use the notation $G \preceq c \cdot  \wt{G}$ as shorthand for
-\begin{equation*}
-x^T L x \leq c \cdot  x^T \wt{L} x,~~\textrm{for all $x \in \Reals^n$}.
-\end{equation*}
-
-Let $G_e = (V, \set{e})$ and $P_e = (V, \set{\widetilde{e}: \widetilde{e} \in \gamma(e)})$ be the graphs associated with $e$ and $\gamma(e)$, respectively. By Lemma \ref{lem: path_poincare}, we have
-\begin{equation*}
-G_{e} \preceq \abs{P_e} P_e
-\end{equation*}
-Summing over all $e \in E_G$, we obtain
-\begin{align*}
-G & \preceq \sum_{e \in E_G} \abs{P_e} P_e \\
-& \preceq \max_{e \in E_G} \abs{\gamma(e)} \sum_{e \in E_G} P_e \\
-& \preceq \max_{e \in E_G} \abs{\gamma(e)} b_{\gamma}\cdot \widetilde{G}
-\end{align*}
-
-\subsubsection{Step 2: Mapping $E_{G_{n,r}}$ to $\mathcal{P}_{\overline{G}}$}
-We assume there exists a mapping $\pi:\overline{X} \to X$ which satisfies \eqref{eqn:slepcev_transport_distance}.
-We will apply the Poincare inequality to $\pi^{-1}(G_{n,r})$, the isomorphism of $G_{n,r}$ induced by the mapping of vertices $\pi^{-1}: X \to \overline{X}$.
-We will map a given $(\overline{x}_{k},\overline{x}_{\ell}) \in \pi^{-1}(E)$ to a shortest path $P \in \mathcal{P}_{\overline{G}}$ (measured by Manhattan distance) between $\overline{x}_k$ and $\overline{x}_{\ell}$.
-Recalling the notation $t = n^{1/d}$ and letting
-\begin{equation*}
-s_i = \frac{1}{t}(\textrm{sign}(k_i - \ell_i),0,\ldots,0),~~\textrm{for $i = 1,\ldots,d$,}
-\end{equation*}
-our mapping $\gamma: \pi^{-1}(E) \to \mathcal{P}_{\overline{G}}$ is given by
-\begin{align*}
-\gamma((\overline{x}_{k},\overline{x}_{\ell})) = \bigl(& (\overline{x}_k, \overline{x}_{k} + s_1), (\overline{x}_{k} + s_1,\overline{x}_{k} + 2s_1),\ldots,(\overline{x}_{k} + (\abs{k_1 - \ell_1} - 1)s_1, \overline{x}_{k} + \frac{1}{t}(k_1 - \ell_1)) \\
-& (\overline{x}_{k} + \frac{1}{t}(k_1 - \ell_1), \overline{x}_{k} + \frac{1}{t}(k_1 - \ell_1) + s_2),\ldots,(\overline{x}_{k} + \frac{1}{t}(k_1 - \ell_1) + (\abs{k_2 - \ell_2} - 1)s_2, \overline{x}_{k}  + \frac{1}{t}((k_1 - \ell_1) + (k_2 - \ell_2)) \\
-& \vdots \\
-& (\overline{x}_{k} + \frac{1}{t}\sum_{i = 1}^{d - 1}(k_i - \ell_i) + s_d,\overline{x}_{k} + \frac{1}{t}\sum_{i = 1}^{d - 1}(k_i - \ell_i) + 2s_d), \ldots, (\overline{x}_k + \frac{1}{t}\sum_{i = 1}^{d - 1}(k_i - \ell_i) + (\abs{k_d - \ell_d} - 1)s_d, \overline{x}_{\ell})\bigr)
-\end{align*}
-As a result of~\eqref{eqn:slepcev_transport_distance} we obtain the following bounds on $M(\gamma)$ and $b(\gamma)$:
-\begin{align}
-M(\gamma) & \leq c n^{1/d} \left(\left(\frac{\log n}{n}\right)^{1/d} + r\right) \label{eqn:maximum_path_length}\\
-b(\gamma) & \leq c \left(n^{1/d} \left(\left(\frac{\log n}{n}\right)^{1/d} + r\right)\right)^{2d}, \label{eqn:bottleneck}
-\end{align}
-which we now prove.
-
-\paragraph{Proof of~\eqref{eqn:maximum_path_length}:}
-
-Suppose $(\overline{x}_k, \overline{x}_{\ell}) \in \pi^{-1}(E)$. Then
-\begin{align*}
-\gamma(\overline{x}_k, \overline{x}_{\ell})) = \norm{\overline{x}_k - \overline{x}_{\ell}}_1 & \leq \sqrt{d} \norm{\overline{x}_k - \overline{x}_{\ell}}_2 \\
-& \leq \sqrt{d}\bigl(\norm{\overline{x}_k - \pi(\ol{x}_{k})}_2 + \norm{\pi(\ol{x}_k) - \pi(\ol{x}_{\ell})}_2 + \norm{\overline{x}_\ell - \pi(\ol{x}_{\ell})}_2\bigr) \\
-& \leq \sqrt{d}\left(\norm{\pi(\ol{x}_k) - \pi(\ol{x}_{\ell})}_2 + 2c\left(\frac{\log n}{n}\right)^{1/d}\right) \tag{by \eqref{eqn:transport_distance}} \\
-& \leq \sqrt{d}\left(r + 2c\left(\frac{\log n}{n}\right)^{1/d}\right) \tag{since $(\pi(\ol{x}_k),\pi(\ol{x}_{\ell})) \in E$}
-\end{align*}
-
-\paragraph{Proof of~\eqref{eqn:bottleneck}:}
-
-Suppose $(\overline{x}_k, \overline{x}_{\ell}) \in \gamma(e)$ for some $e = (\overline{x}_i, \overline{x}_j) \in \pi^{-1}(E)$.  Then, by the construction of $\gamma$,
-\begin{equation}
-\label{eqn:bottleneck_pf_1}
-\norm{\overline{x}_k - \overline{x}_i}_2 \wedge \norm{\overline{x}_\ell - \overline{x}_j}_2  \leq \norm{\overline{x}_j - \overline{x}_i}_2
-\end{equation}
-As shown in the preceding paragraph, assuming~\eqref{eqn:transport_distance} we have
-\begin{equation*}
-\norm{\overline{x}_j - \overline{x}_i}_2 \leq r + 2c\left(\frac{\log n}{n}\right)^{1/d}
-\end{equation*}
-and therefore by~\eqref{eqn:bottleneck_pf_1},
-\begin{align*}
-\abs{\set{e \in E: (\overline{x}_k, \overline{x}_{\ell}) \in \gamma(e)}} & \leq \abs{\set{\overline{x}_i: \norm{\overline{x}_k - \overline{x}_i}_2 \leq r + 2c\left(\frac{\log n}{n}\right)^{1/d}}} \cdot \abs{\set{\overline{x}_i: \norm{\overline{x}_k - \overline{x}_i}_2 \leq r + 2c\left(\frac{\log n}{n}\right)^{1/d}}} \\
-& \leq \left(n^{1/d}\left( r + 2c\left(\frac{\log n}{n}\right)^{1/d}\right)\right)^2.
-\end{align*}
-
-Putting the pieces together, we have that when $d = 1$ and $r = (\log n)^a (\log n/n)$, then
-\begin{equation*}
-G_{n,r} \preceq c \cdot (\log n)^{(1/d + a) + 2 + 2da}\cdot \overline{G}
-\end{equation*}
-with probability at least $1 - o(1)$.
-
-\subsection{Proof of Lemma~\ref{lem:pi_max_pf_1}}
-
-Write $\overline{L} = \overline{V}\overline{S} \overline{V}^T$ for the spectral decomposition of $\overline{L}$. The proof of Lemma~\ref{lem:pi_max_pf_1} will proceed according to the following steps.
-\begin{enumerate}
-	\item We show that the incoherence $\max_{i} v_{k,i}^2$ can be upper bounded
-	\begin{equation}
-	\label{eqn:incoherence_proof_1}
-	\max_{i} v_{k,i}^2 \leq \frac{1}{n} \left(\sum_{j = 1}^{n} \abs{\dotp{v_k}{\overline{v}_j}}\right)^2
-	\end{equation}
-	replacing the norm $\norm{v_k^2}_{\infty}$ by the inner products $\dotp{v_k}{\overline{v}_j}$.
-	\item Let $I(k,r) = \set{j \geq 0: \abs{j - k} \leq r}$. Using Davis-Kahan, we establish that the following upper bounds
-	\begin{equation}
-	\label{eqn:incoherence_proof_2}
-	\sum_{j \not\in I(k,r)}^{n} (\dotp{v_k}{\overline{v}_j})^2 \leq 400 \frac{n^4 \delta^2 s_{\max}^2}{R^4}
-	\end{equation}
-	hold for each $k \leq n/8$ and any $r \geq 8k\sqrt{\delta}$. 
-	\item We carefully upper bound the $L_1$ norm present in \eqref{eqn:incoherence_proof_1} given the various bounds on $L_2$ norm we've established in \eqref{eqn:incoherence_proof_2}.
-\end{enumerate}
-
-\paragraph{Step 1: Proof of~\eqref{eqn:incoherence_proof_1}.}
-
-We can re-express $v_k$ in the basis $(\overline{v}_j)$ as
-\begin{equation*}
-v_k = \sum_{j = 1}^{n} \dotp{v_k}{\overline{v}_j} \overline{v}_j,
-\end{equation*}
-whereupon the bound~\eqref{eqn:incoherence_proof_1} follows from the incoherence property of the chain
-\begin{equation*}
-\max_{i,k = 1,\ldots,n} \overline{v}_{k,i}^2 \leq \frac{1}{n}.
-\end{equation*}
-\paragraph{Step 2: Proof of~\eqref{eqn:incoherence_proof_2}.}
-
-Let $\widetilde{L} = L + H$. By Davis Kahan, we have that
-\begin{equation}
-\label{eqn:davis_kahan}
-\sum_{j \not\in I(k,r)} (\dotp{v_k}{\overline{v}_j})^2 \leq \left(\frac{\norm{H}_{\textrm{op}}}{\min{\abs{\overline{s}_j - s_k}:j \in I(k,r)}}\right)^2
-\end{equation}
-To upper bound the numerator, we use~\eqref{eqn:spectral_similarity} to get
-\begin{equation*}
-\norm{H}_{op} \leq \delta s_{n}.
-\end{equation*}
-
-To lower bound the denominator, we note
-\begin{align}
-\abs{\overline{s}_j - s_k} & \geq \abs{\overline{s}_j - \overline{s}_k} - \abs{\overline{s}_k - s_k} \nonumber \\
-& \geq \abs{\overline{s}_j - \overline{s}_k} - \delta \overline{s}_k. \label{eqn:incoherence_proof_8}
-\end{align}
-The eigenvalues of the chain graph are well known to be
-\begin{equation*}
-\overline{s}_j = 2\left(1 - \cos\left(\frac{j\pi}{n}\right)\right) = 4\sin^2\left(\frac{k\pi}{n}\right) ~~\textrm{for $k = 0,\ldots,n - 1$.}
-\end{equation*}
-By Taylor expansion, for $k \leq j \leq n/(2\pi)$ we have
-\begin{equation}
-\label{eqn:incoherence_proof_5}
-\overline{s}_j - \overline{s}_k = 2\left(\cos\left(\frac{k\pi}{n}\right) - \cos\left(\frac{j\pi}{n}\right)\right) \geq \frac{(k - j)^2\pi^2}{n^2}.
-\end{equation}
-and when $j \geq n/(2\pi)$ we have $\overline{s}_j - \overline{s}_k \geq 2(\cos\left(\frac{\pi}{8}\right) - \cos\left(\frac{1}{2}\right) > .09$.
-Additionally since $\sin(x) \leq x$ for all $x \geq 0$, we have
-\begin{equation}
-\label{eqn:incoherence_proof_6}
-\overline{s}_k \leq 4\frac{k^2\pi^2}{n^2}.
-\end{equation}
-Combining~\eqref{eqn:incoherence_proof_5},\eqref{eqn:incoherence_proof_6}, and the lower bound $\abs{k - j} \geq R \geq 8k\sqrt{\delta}$, we have that
-\begin{equation}
-\label{eqn:incoherence_proof_7}
-\abs{\overline{s}_j - \overline{s}_k} - \delta \overline{s}_k \geq .04\frac{(k - j)^2\pi^2}{n^2} - 4 \delta\frac{k^2\pi^2}{n^2} \geq .02\frac{(k - j)^2\pi^2}{n^2}.
-\end{equation}
-The result follows from \eqref{eqn:incoherence_proof_7}, \eqref{eqn:incoherence_proof_8} and \eqref{eqn:davis_kahan}.
-
-\paragraph{Step 3: $L_1$ norm to $L_2$ norm.}
-
-Let $u \in \Reals^n$, and suppose we know that the $L_2$ norm of $u$ is bounded,
-\begin{equation}
-\label{eqn:incoherence_proof_3}
-\norm{u}_2^2 \leq 1.
-\end{equation}
-Under no other conditions on $u$, the upper bound $\norm{u}_1 \leq \sqrt{n}$ is the best that can be hoped for (achieved when $u = (n^{-1/2},\ldots,n^{-1/2})$).
-However, suppose we also know that for some $B_2^2 \geq B_3^2 \geq \ldots \geq B_n^2$, we have that there exists some $R \in [n]$ such that
-\begin{equation}
-\label{eqn:incoherence_proof_4}
-\sum_{j = r}^{n} v_j^2 \leq B_r^2 ~~\textrm{for each $r = R,\ldots,n$.}
-\end{equation}
-Clearly, if $B_n^2 \leq \frac{1}{n}$ then $u$ cannot be equal to $(n^{-1/2},\ldots,n^{-1/2})$.
-We might hope for more general improvements on the bound $\norm{u}_1 \leq \sqrt{n}$ if $B_R^2$ is quite small once $R$ gets sufficiently large.
-The following Lemma gives such a result.
-\begin{lemma}
-	\label{lem:pi_max_pf_1_util_1}
-	Suppose $u \in \Reals^n$ satisfies \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4} for some $1 = B_1^2 = B_2^2 \geq \ldots \geq B_n^2 \geq B_{n+1}^2 = 0$.
-Assume additionally that there exists an $R \in [n]$ such that
-	\begin{equation}
-	\label{eqn:incoherence_util_1}
-	B_r^2 - B_{r + 1}^2 \leq \frac{1}{r}~\textrm{for each $r = R, R+1,\ldots, n - 1$.}
-	\end{equation}
-	Then,
-	\begin{equation}
-	\label{eqn:incoherence_util_2}
-	\norm{u}_1 \leq \sqrt{R}\sqrt{1 - B_R^2} + \sum_{j = R}^{n} \sqrt{B_j^2 - B_{j+1}^2}.
-	\end{equation}
-\end{lemma}
-\begin{proof}
-	Set
-	\begin{equation*}
-	(u_\star)_j^2 = 
-	\begin{cases*}
-	\frac{1}{R}(1 - B_R^2), ~ j \leq R \\
-	B_j^2 - B_{j + 1}^2, j > R
-	\end{cases*}
-	\end{equation*}
-	so that $\norm{u_\star}_1$ is equal to the right hand side of~\eqref{eqn:incoherence_util_2}.
-We now prove by contradiction that $u_\star$ maximizes $\norm{\cdot}_1$ under the constraints \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4}.
-Suppose this were not true, and that the maximum is achieved for some $u \neq u_{\star}$.
-	\begin{itemize}
-		\item since $\norm{u}_1 > \norm{u_{\star}}$ then there exists some index $j$ such that
-		\begin{equation}
-		\label{eqn:incoherence_util_3}
-		u_j > (u_{\star})_j
-		\end{equation} 
-		\item by~\eqref{eqn:incoherence_proof_3} $\norm{u}_2 \leq 1$. Since $\norm{u_{\star}}_2 = 1$, by \eqref{eqn:incoherence_util_3} there must exist some index $k$ such that
-		\begin{equation}
-		\label{eqn:incoherence_util_4}
-		(u_{\star})_k > u_k.
-		\end{equation} 
-		Choose $k$ to be the largest of all such indices.
-		\item suppose $k < j$ and $j > R$.
-Since $k$ was chosen to be the largest such index which satisfied~\eqref{eqn:incoherence_util_4}, clearly $v_i \geq (u_{\star})_i$ for all $i > j$, and by \eqref{eqn:incoherence_util_3} $u_j > (u_{\star})_j$.
-But then
-		\begin{equation*}
-		B_j^2 = \sum_{i = j}^{N} (u_{\star})_i^2 < \sum_{i = j}^{N} u_i^2 
-		\end{equation*}
-		and so $v$ does not satisfy~\eqref{eqn:incoherence_proof_4}. Therefore either $k > j$ or $j \leq R$.
-		\item In either case, we have
-		\begin{equation*}
-		u_j > (u_{\star})_j \geq (u_{\star})_k > u_k.
-		\end{equation*}
-		Let $j \leq l < k$ be the largest index for which $v_j > (a_{\star})_j$.
-
-		We now construct a vector $\wt{u}$ which satisfies the constraints \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4} such that $\norm{\wt{u}}_1 > \norm{u}_1$.
-Once we have shown this, we will have established a contradiction, and the proof of Lemma~\ref{lem:pi_max_pf_1_util_1} will be complete.
-Let $\wt{u} = (\wt{u}_i)$ be given as follows:
-		\begin{equation*}
-		\wt{u}_i = 
-		\begin{cases*}
-		u_i, ~~\textrm{if $i \neq k,l$}, \\
-		u_{\star,k}, ~~\textrm{if $i = k$}, \\
-		\sqrt{u_l^2 - (u_{\star,k}^2 - a_k^2)}, ~~\textrm{if $i = l$.}
-		\end{cases*}
-		\end{equation*}
-		Clearly $\norm{\wt{u}}_2 = \norm{u}_2 = 1$ and so $\wt{u}$ satisfies~\eqref{eqn:incoherence_proof_3}.
-To see that $\wt{u}$ satisfies~\eqref{eqn:incoherence_proof_4} we separate the analysis into cases.
-When $r \geq l + 1$, we have that
-		\begin{equation*}
-		\sum_{i = r}^{n} \wt{u}_i^2 = \sum_{i = r}^{n} u_i^2 \leq B_r^2.
-		\end{equation*}
-		When $l < r \leq k$, we have that
-		\begin{equation*}
-		\sum_{i = r}^{n} \wt{u}_i^2 \leq \sum_{i = r}^{k} u_{\star,i}^2 + \sum_{i = k +1}^{n} u_{i}^2 \leq B_r^2 - B_{k + 1}^2 + B_{k + 1}^2 = B_r^2
-		\end{equation*}
-		Finally, when $R \leq r \leq l$, we have
-		\begin{equation*}
-		\sum_{i = r}^{n} \wt{u}_i^2 \leq u_l^2 - (u_{\star,k}^2 - u_k^2) + (u_{\star,k}^2) + \sum_{i \neq k,l} u_i^2 = \sum_{i = r}^{n} u_i^2 \leq B_r^2.
-		\end{equation*}
-		Therefore $\wt{u}$ satisfies~\eqref{eqn:incoherence_proof_4}. Finally, we have that
-		\begin{equation*}
-		\norm{\wt{u}}_1 = \norm{u}_1 + (u_{\star,k}  - u_k) - (\sqrt{u_l^2 - (u_{\star,k}^2 - u_k^2)} - u_l) > \norm{u}_1 + (u_{\star,k}  - u_k) - (u_{\star,k}  - u_k) = \norm{u}_1,
-		\end{equation*}
-		so we have established the desired contradiction.
-	\end{itemize}
-\end{proof}
-
-\paragraph{Putting the pieces together.}
-We apply Lemma~\ref{lem:pi_max_pf_1_util_1} to the vector $u = (u_i)_{i = 1}^{n}$, where
-\begin{equation*}
-u_i = \frac{1}{\sqrt{2}}\left(\abs{\dotp{v_k}{\overline{v}_{k - i}}} + \abs{\dotp{v_k}{\overline{v}_{k + i}}}\right).
-\end{equation*}
-Note the following:
-\begin{itemize}
-	\item $\norm{u}_2 \leq 1$.
-	\item By \eqref{eqn:incoherence_proof_2}, the vector $u$ satisfies the constraint~\eqref{eqn:incoherence_proof_4} with
-	\begin{equation*}
-	B_r^2 = 400\frac{n^4\delta^2s_{n}^2}{r^4} ~~\textrm{when $r \geq 8 k \sqrt{\delta}$.}
-	\end{equation*}
-	\item Taylor expanding $f(r + 1) = (r + 1)^{-4}$ around $r$, we have that
-	\begin{equation*}
-	B_r^2 - B_{r + 1}^2 \leq 3200 \frac{n^4 \delta^2 s_{n}^2}{r^5} ~~\textrm{when $r \geq 2$.}
-	\end{equation*}
-	Therefore, the sequence $(B_r)$ satisfies \eqref{eqn:incoherence_util_1} for all $r$ large enough such that 
-	\begin{equation*}
-	r \geq 2^{5/4} 100^{1/4} n \sqrt{\delta s_{n}}
-	\end{equation*}
-	\item Since $u$ and $(B_r)$ satisfy the constraints \eqref{eqn:incoherence_proof_3}, \eqref{eqn:incoherence_proof_4} and \eqref{eqn:incoherence_proof_5}, we may apply Lemma~\ref{lem:pi_max_pf_1_util_1} to $u$, obtaining that
-	\begin{equation*}
-	\sum_{j = 1}^{n} \abs{\dotp{v_j}{\overline{v}_k}} = \sqrt{2}\norm{u}_1 \leq \sqrt{2R} + 60\sqrt{2} n^2 \delta s_{n} \sum_{r = R}^{n} r^{-5/2}
-	\end{equation*}
-	for any $R \geq 2 \vee k\sqrt{\delta} \vee 2^{5/4} 100^{1/4} n \sqrt{\delta s_{n}}$. Bounding sum by integral, we have that if $R \geq 2$ then
-	\begin{equation*}
-	\sum_{r = R}^{n} r^{-5/2} \leq \int_{R - 1}^{n - 1} x^{-5/2} \,dx \leq \frac{2^{3/2}2}{3} R^{-3/2}.
-	\end{equation*}
-	Plugging this into the previous expression, we arrive at
-	\begin{equation*}
-	\sum_{j = 1}^{n} \abs{\dotp{\overline{v}_j}{v_k}} \leq \sqrt{R} + 60 \sqrt{2} n^2 \delta s_{\max} \frac{2^{3/2}2}{3R^{3/2}}
-	\end{equation*}
-	for any $R \geq 2 \vee k\sqrt{\delta} \vee 2^{5/4} 100^{1/4} n \sqrt{\delta s_{\max}}$. Lemma~\ref{lem:pi_max_pf_1} then follows from~\eqref{eqn:incoherence_proof_1}.
-\end{itemize}
-
-\section{Auxiliary Results}
-
-\subsection{Results of Others}
-
-Let $\overline{X}$ denote the $n$ evenly spaced grid points on $[0,1]^d$; formally, letting $t = n^{1/d}$
-\begin{equation*}
-\overline{X} = \biggl\{\frac{1}{t}(k_1,\ldots,k_d): k \in [t]^d\biggr\}
-\end{equation*}
-
-\begin{theorem}[Theorem 1 of \textcolor{red}{Garcia-Trillos and Slepcev}]
-	\label{thm:slepcev_transport_distance}
-	With probability at least $1 - c/n$ there exists a bijection $\pi: \overline{X} \to X$ such that
-	\begin{equation}
-	\label{eqn:slepcev_transport_distance}
-	\max_{k \in [t]^d} \abs{\overline{x}_k - \pi(\overline{x}_k)} \leq c \left(\frac{\log n}{n}\right)^{1/d}
-	\end{equation}
-\end{theorem}
-
-\begin{theorem}[\textcolor{red}{Evans} Chapter 5.4, Theorem 1]
-	\label{thm:evans_extension}
-	Assume $U$ is bounded and $\partial U$ is $C^1$.
-Select a bounded open set $V$ such that $U \subset \subset V$ ($U$ is compactly contained in $V$).
-Then there exists a bounded linear operator $E: H^1(U) \to H^1(\Rd)$ such that for each $u \in H^1(U)$:
-	\begin{enumerate}
-		\item $Eu = u$ a.e. in $U$,
-		\item $Eu$ has support within $V$, and 
-		\item 
-		\begin{equation*}
-		\norm{Eu}_{H^1(\Rd)} \leq C \norm{u}_{H^1(\Rd)}
-		\end{equation*}
-		the constant $C$ depending only on $U$ and $V$.
-	\end{enumerate}
-\end{theorem}
-
-\begin{lemma}[Poincare inequality for path graphs.]
-	\label{lem: path_poincare}
-	Fix $m \geq 0$. For vertices $V = \set{1, \ldots,m}$ define the path $P(1 \to m) = ((1,2),(2,3),\ldots, (m-1,m))$ and $G_{(1,m)}$ to be the graph consisting only of an edge between $1$ and $m$. Then,
-	\begin{equation*}
-	(m - 1) \cdot P(1 \to m) \succeq G_{(1,m)}
-	\end{equation*}
-\end{lemma}
-
-\subsection{Integrals}
-
-For Lemmas~\ref{lem:expected_first_order_seminorm} - \ref{lem:expected_first_order_seminorm_3}, we will assume that $K$ is a kernel function compactly supported on $B(0,1)$ and upper bound $K(x) \leq K_{\max}$.
-Additionally, we note that although each Lemma assumes $g \in C^1(\Xset)$, in each case if we assume merely $g \in H^1(\Xset)$ the results still hold due to the density of $C^{\infty}$ in $H^1(\Xset)$.
-
-\begin{lemma}
-	\label{lem:expected_first_order_seminorm}
-	Suppose $g \in C^{1}(\Xset)$ for $\Xset$ a Lipschitz domain and that $\abs{p(x)} \leq p_{\max}$ for all $x \in \Xset$. Then
-	\begin{equation*}
-	\Ebb\Bigl[(g(x_j) - g(x_i))^2K_r(x_i,x_j)\Bigr] \leq c K_{\max} p_{\max}^2 r^2 [g]_{H^1(\Xset)}^2
-	\end{equation*}
-	for a constant $c$ which depends only on $\Xset$ and $d$.
-\end{lemma}
-\begin{proof}
-	Since $\Xset$ is a Lipschitz domain, we may take $h \in C^1(\Rd)$ to be an extension of $g$ such that $h = g$ a.e. on $\Xset$, and additionally
-	\begin{equation*}
-	[h]_{H^1(\Rd)} \leq c[g]_{H^1(\Xset)}
-	\end{equation*}
-	Since $h = g$ a.e on $\Xset$, the expectations satisfy
-	\begin{equation*}
-	\Ebb\Bigl[(g(x_j) - g(x_i))^2K_r(x_i,x_j)\Bigr] = \Ebb\Bigl[(h(x_j) - h(x_i))^2K_r(x_i,x_j)\Bigr].
-	\end{equation*}
-	By the fundamental theorem of calculus we have for any $y,x \in \Rd$,
-	\begin{equation}
-	\label{eqn:expected_first_order_seminorm_pf1}
-	h(y) - h(x) = \int_{0}^{1} \frac{d}{dt}\bigl[h(x + t(y - x))\bigr] \,dt = \int_{0}^{1} \dotp{\nabla(h(x + t(y - x)))}{y - x} \,dt
-	\end{equation}
-	where the integral is well-defined as $\nabla h(z)$ exists almost everywhere since $h \in C^1(\Rd)$. We now perform some standard calculus:
-	\begin{align*}
-	\Ebb[(h(x_j) - h(x_i))^2K_r(x_i,x_j)] & \leq p_{\max}^2 \int_{\Rd} \int_{\Rd} (h(y) - h(x))^2 K_r(y,x) \,dy \,dx\\
-	& = p_{\max}^2 \int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \dotp{\nabla(h(x + t(y - x)))}{y - x} \,dt\right)^2 K_r(y,x) \,dy \,dx \\
-	& \overset{(i)}{\leq} p_{\max}^2 \int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \norm{\nabla(h(x + t(y - x)))}\norm{y - x} \,dt\right)^2 K_r(y,x) \,dy \,dx \\
-	& \overset{(ii)}{\leq} p_{\max}^2 r^2\int_{\Rd} \int_{\Rd} \left(\int_{0}^{1} \norm{\nabla(h(x + t(y - x)))}\,dt\right)^2 K_r(y,x) \,dy \,dx \\
-	& \overset{(iii)}{\leq} p_{\max}^2 r^2\int_{\Rd} \int_{\Rd} \int_{0}^{1} \norm{\nabla(h(x + t(y - x)))}^2 \,dt K_r(y,x) \,dy \,dx \\
-	& \overset{(iv)}{\leq} p_{\max}^2 K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(x,r)} \norm{\nabla(h(x + t(y - x)))}^2 \,dy \,dt \,dx \\
-	& \overset{(v)}{\leq}  p_{\max}^2 K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(0,r)} \norm{\nabla(h(x + z))}^2  \,dz \,dt \,dx
-	\end{align*}
-	where $(i)$ follows by Cauchy-Schwarz, $(ii)$ follows since either $\norm{y - x} \leq r$ or $K_r(y,x) = 0$, $(iii)$ follows by Jensen's, $(iv)$ follows by the assumption $K \leq K_{\max}$ supported on $B(0,1)$, and $(v)$ follows from the change of variables $z = x + t(y - x)$.
-Finally, again using Fubini's Theorem, we have
-	\begin{align*}
-	K_{\max} r^{2 - d}\int_{\Rd} \int_{0}^{1} \int_{B(0,r)} \norm{\nabla(h(x + z))}^2  \,dz \,dt \,dx & = r^{2 - d}\int_{B(0,r)} \int_{0}^{1} \int_{\Rd} \norm{\nabla(h(x + z))}^2  \,dz \,dt \,dx \\
-	& = K_{\max} r^2 [h]_{W_d^{1,2}(\Rd)}.
-	\end{align*}
-\end{proof}
-
-\begin{lemma}
-	\label{lem:expected_first_order_seminorm_2}
-	Suppose $f \in C^{1}(\Xset)$ for $\Xset$ a Lipschitz domain and that $\abs{p(x)} \leq p_{\max}$ for all $x \in \Xset$. Then
-	\begin{equation*}
-	\Ebb\Bigl[\abs{D_if(x_h)}\cdot\abs{D_if(x_j)} \Bigr] \leq c K_{\max}^3 p_{\max}^3 r^2 [f]_{H^1(\Xset)}^2
-	\end{equation*}
-	for a constant $c$ which depends only on $\Xset$ and $d$.
-\end{lemma}
-\begin{proof}
-	Since $\Xset$ is a Lipschitz domain, we may take $g \in C^1(\Rd)$ to be an extension of $f$ such that $g = f$ a.e. on $\Xset$, and additionally
-	\begin{equation*}
-	[g]_{H^1(\Rd)} \leq c[f]_{H^1(\Xset)}
-	\end{equation*}
-	Since $g = f$ a.e on $\Xset$, the expectations satisfy
-	\begin{equation*}
-	\Ebb\Bigl[\abs{D_if(x_h)}\cdot\abs{D_if(x_j)} \Bigr] =\Ebb\Bigl[\abs{D_ig(x_h)}\cdot\abs{D_ig(x_j)} \Bigr]
-	\end{equation*}
-
-	We rewrite $\Ebb\bigl[\abs{D_{i}g(x_j)} \cdot \abs{D_ig(x_h)} \bigr]$ as follows,
-	\begin{align*}
-	\Ebb\Bigl[\abs{D_{i}g(x_j)} \cdot \abs{D_ig(x_h)} \Bigr] & = \int \int \int \abs{g(z) - g(x)} \cdot \abs{g(z) - g(y)} K_r(z,y) K_r(z,x) \,dP(x) \,dP(y) \,dP(x) \\
-	& = \int \left[\int \abs{g(z) - g(x)} K_r(z,x) \,dP(x)\right]^2 \,dP(z) \\
-	& \leq p_{\max}^3 \int_{\Xset} \left[\int_{\Xset} \abs{g(z) - g(x)} K_r(z,x) \,dx\right]^2 \,dz
-	\end{align*}
-	Applying~\eqref{eqn:expected_first_order_seminorm_pf1} inside the integral gives
-	\begin{align*}
-	\int_{\Xset} \abs{g(z) - g(x)} K_r(z,x) \,dx & \leq \int_{\Rd} \abs{g(z) - g(x)} K_r(z,x) \,dx \\
-	& = \int_{\Rd} \abs{\int_{0}^{1} \dotp{\nabla g(x + t(z - x))}{z - x} \,dt} K_r(z,x) \,dx \\
-	& \leq \int_{\Rd} \int_{0}^{1} \norm{\nabla g(x + t(z - x))}\cdot\norm{z - x} \,dt K_r(z,x) \,dx \\
-	& \leq r \int_{\Rd} \int_{0}^{1} \norm{\nabla g(x + t(z - x))} \,dt K_r(z,x) \,dx \\
-	& \leq r \frac{K_{\max}}{r^d} \int_{B(z,r)} \int_{0}^{1} \norm{\nabla g(x + t(z - x))} \,dt  \,dx \\
-	& \leq r K_{\max} \int_{B(0,1)} \int_{0}^{1} \norm{\nabla g(x - try)} \,dt  \,dy,
-	\end{align*}
-	and as a result, 
-	\begin{equation*}
-	p_{\max}^3 \int_{\Xset} \left[\int_{\Xset} \abs{g(z) - g(x)} K_r(z,x) \,dx\right]^2 \,dz \leq c\cdot p_{\max}^3 r^2 K_{\max}^3 [f]_{W_d^{1,2}(\Rd)}^2.
-	\end{equation*}
-\end{proof}
-
-\begin{lemma}
-	\label{lem:expected_first_order_seminorm_3}
-	Suppose $f \in C^{1}(\Xset)$ and that $\abs{p(x)} \leq p_{\max}$ for all $x \in \Xset$. Then for any distinct $i,j,k,\ell$ each in $[n]$,
-	\begin{equation*}
-	\Ebb\Bigl[\abs{D_{k}f(x_i)}\cdot{\abs{D_{\ell}f(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr] \leq c K_{\max} p_{\max}^2 r^{2 + d} [f]_{H^1(\Rd)}^2
-	\end{equation*}
-	for a constant $c$ which depends only on $\Xset$ and $d$.
-\end{lemma}
-\begin{proof}
-	Since $\Xset$ is a Lipschitz domain, we may take $g \in C^1(\Rd)$ to be an extension of $f$ such that $g = f$ a.e. on $\Xset$, and additionally
-	\begin{equation*}
-	[g]_{H^1(\Rd)} \leq c[f]_{H^1(\Xset)}
-	\end{equation*}
-	Since $g = f$ a.e on $\Xset$, the expectations satisfy
-	\begin{equation*}
-	\Ebb\Bigl[\abs{D_{k}f(x_i)}\cdot{\abs{D_{\ell}f(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr] = 	\Ebb\Bigl[\abs{D_{k}g(x_i)}\cdot{\abs{D_{\ell}g(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr].
-	\end{equation*}
-	We rewrite the expectation as an integral,
-	\begin{align*}
-	\Ebb\Bigl[& \abs{D_{k}g(x_i)}\cdot{\abs{D_{\ell}g(x_j)}}\cdot\1\{\norm{x_i - x_j} \leq (2q + 1)r\}\Bigr] \\
-	& \leq p_{\max}^4 \int_{\Xset^4} \abs{g(x) - g(y)} \cdot \abs{g(u) - g(v)} \cdot  K_r(x,y) K_r(u,v) \1\{\norm{y - v} \leq (2q + 1)r\} \,dy \,dx \,du \,dv
-	\end{align*}
-	By substituting $z_1 = (y - v)/r$, $z_2 = (u - v)/r$, and $z_3 = (x - y)/r = (x - v)/r + z_1$, we can simplify the integral in the previous display,
-	\begin{align*}
-	\int_{\Xset^4} & \abs{g(x) - g(y)} \cdot \abs{g(u) - g(v)} \cdot K_r(x,y) K_r(u,v) \1\{\norm{y - v} \leq (2q + 1)r\} \,dy \,dx \,du \,dv \\
-	& \leq K_{\max}^2 r^d \int_{\Xset} \int_{[B(0,1)]^3} \abs{g\bigl((z_3 + z_1)r + v\bigr) - g(z_1r + v)} \cdot \bigl|g(z_2r + v) - g(v)\bigr| \,dz_1 \,dz_2 \,dz_3 \,dv \\
-	& \leq  K_{\max}^2 r^{d + 2} \int_{[B(0,1)]^3} \int_{[0,1]^2} \int_{\Xset} \norm{\nabla g(t z_3 r + z_1r + v)} \cdot \norm{\nabla g(t z_2 r + v)} \,dv \,dt_1 \,dt_2 \,dz_1 \,dz_2 \,dz_3 \\
-	& \leq c \nu_d^3 K_{\max}^2 r^{d + 2} [f]_{W_{d}^{1,2}(\Xset)}^2.
-	\end{align*}
-\end{proof}
-
-\begin{lemma}
-	\label{lem:remainder_term}
-	Suppose that $f \in \Leb^2(U)$ for some open domain $U$, and that $h:U \times U \times [0,1] \to \Reals$ is uniformly bounded.
-Then, the function $g(x) = \int_{0}^{1} \int_{B(0,1)} f(x + aty) h(y,x, t) \,dy \,dt$ also belongs to $\Leb^2(U)$, with norm
-	\begin{equation*}
-	\norm{g}_{\Leb^2(U)} \leq \nu_d \cdot \norm{f}_{\Leb^2(U)} \cdot \norm{h}_{\infty}
-	\end{equation*}
-\end{lemma}
-\begin{proof}
-	We compute the squared norm of $g$,
-	\begin{align*}
-	\norm{g}_{\Leb^2(\Rd)}^2 & = \int_{U} \left(\int_{0}^{1} \int_{B(0,1)} f(x + aty) h(y,x,t) \,dt \,dy \right)^2 \,dx \\
-	& \leq \norm{h}_{\infty}^2 \int_{U} \left(\int_{0}^{1} \int_{B(0,1)} f(x + aty) \,dt \,dy \right)^2 \,dx \\
-	& \leq \nu_d^2 \norm{h}_{\infty}^2 \int_{U} \int_{0}^{1} \frac{1}{\nu_d}\int_{B(0,1)} f^2(x + aty) \,dt \,dy \,dx \tag{Jensen's inequality} \\
-	& = \nu_d^2 \norm{h}_{\infty}^2 \int_{0}^{1} \int_{B(0,1)} \frac{1}{\nu_d}\int_{U}f^2(x + aty) \,dt \,dy \,dx \tag{Fubini's theorem} \\
-	& = \nu_d^2 \norm{h}_{\infty}^2 \norm{f}_{\Leb^2(U)}^2.
-	\end{align*}
-\end{proof}
 
@@
 \end{lemma}
-\begin{proof}
+
+\section{Proofs}
+
+\subsection{Proof of Lemma~\ref{lem:leading_term_sobolev_compact}}
+
 One can interpret the conclusions of Lemma~\ref{lem:leading_term_sobolev_compact} as demonstrating that expected difference operators behave similarly to derivatives over $H_0^{s}(\Xset)$.
@@
 \end{align*}
-	and the statement follows by Lemma~\ref{lem:remainder_term}. The same result holds (up to different constants) for $s = 0$ and general $q$ by induction.
+and the statement follows by Lemma~\textcolor{red}{11}. The same result holds (up to different constants) for $s = 0$ and general $q$ by induction.
 
@@
 \end{equation*}
-	thus establishing~\eqref{eqn:leading_term_sobolev_compact_2} when $s = 2$. Otherwise when $s > 2$, we rearrange
+thus establishing~\eqref{eqn:leading_term_holder_compact_2} when $s = 2$. Otherwise when $s > 2$, we rearrange
 \begin{equation*}
@@
 \end{equation}
-	establishing~\eqref{eqn:leading_term_sobolev_compact_1} when $s > 2$.
+establishing~\eqref{eqn:leading_term_holder_compact_1} when $s > 2$.
 
@@
 Rewriting the double sum as a single sum over $\ell + m = 2q,\ldots,s - 1$ and plugging back in to~\eqref{eqn:leading_term_sobolev_compact_pf6} gives the desired result~\eqref{eqn:leading_term_sobolev_compact_4}.
+
+\subsection{Proof of Lemma~\ref{lem:leading_term_holder_compact}}
+
+The proof of Lemma~\ref{lem:leading_term_holder_compact} is centered on taking Taylor expansions of the function $f$ and the density $p$.
+When $s \geq 1$, since $f \in C_c^s(\Xset)$ it admits a Taylor expansion of the form
+\begin{equation*}
+f(y) =  \sum_{\abs{\alpha} = 0}^{s - 1} f^{(\alpha)}(x) (z - x)^{\alpha} + \sum_{\abs{\alpha} = s} (y - x)^{\alpha} F_{\alpha,x}(y),
+\end{equation*}
+and when $s \geq 2$, since $p \in C_c^{s - 1}(\Xset)$ it also admits a Taylor expansion,
+\begin{equation*}
+p(y) = \sum_{\abs{\beta} = 0}^{s - 2} p^{\beta}(x) (y - x)^{\beta} + \sum_{\abs{\beta} = s} (x - y)^s G_{\beta}(x,y).
+\end{equation*}
+In both cases $x$ and $y$ are arbitrary points in $\Xset$, and we use the integral form of the remainders:
+\begin{align*}
+F_{\alpha}(x,y) & = \int_{0}^{1} f^{(\alpha)}(x + t(y - x)) \,dt, && \Bigl|F_{\alpha}(x,y)\Bigr| \leq \norm{f}_{C^s(\Xset)} \\
+G_{\beta}(x,y) & = \int_{0}^{1} p^{(\beta)}(x + t(y - x)) \,dt, && \Bigl|G_{\beta}(x,y)\Bigr| \leq \norm{p}_{C^{s - 1}(\Xset)}
+\end{align*}
+where by Rademacher's Theorem $f^{(\alpha)}$ and $p^{(\beta)}$ exist almost everywhere, and the preceding integrals are therefore well defined.
+
+It will also be helpful to introduce some notation. For $F: \Xset \times \Xset \to \Reals$, let
+\begin{align*}
+\Bigl(\Ebb_{\alpha}[F]\Bigr)(x) & := \int (y - x)^{\alpha} F(x,y) K_r(y,x) p(y) \,dy,~~ && \Ebb_{\alpha,x} := \Bigl(\Ebb_{\alpha}[1]\Bigr)(x) \\
+\Bigl(\Ibb_{\alpha}[F]\Bigr)(x) & := \int (y - x)^{\alpha} F(x,y) K_r(y,x) \,dy,~~ && \Ibb_{\alpha,x}  := \Bigl(\Ibb_{\alpha}[1]\Bigr)(x)
+\end{align*}
+Additionally we define
+\begin{equation*}
+I_{\alpha} := \int z^{\alpha} K\bigl(\norm{z}\bigr) \,dz.
+\end{equation*}
+and note that when $B(x,r) \subset \Xset$, the following two facts are true: first, that $\Ibb_{\alpha,x} = r^{\abs{\alpha}} I_{\alpha}$, and second that $I_{\alpha} = 0$ when $\abs{\alpha} = 1$.
+
+The case when $s = 0$ follows from the boundedness of $f$, as a simple inductive argument yields
+\begin{equation*}
+\abs{\Ebb\Bigl[D_kf(x)\Bigr]} \leq 2^qp_{\max}^q \norm{f}_{C^0(\Xset)}.
+\end{equation*}
+for any $x \in \Xset$.
+
+For $s \geq 1$, we first show the desired estimate when $x \in \Xset \setminus \Xset_{qr}$.
+
+\subsubsection{Boundary point.}
+
+Since $f \in C_c^{s}(\Xset)$, there exists an open set $V$ compactly contained $V \subset \bar{V} \subset U$ such that $\mathrm{supp}(f) \subset V$.
+Furthermore, since $\dist(x,\partial \Xset) < r$ there exists a particular $x_0$ in $\Xset \setminus V$ such that $\norm{x_0 - x} \leq r$.
+Since there exists some $\delta > 0$ such that $B(x_0,\delta) \subset \Xset \setminus V$, we know
+\begin{equation*}
+f^{(\beta)}(x_0) = 0,~~\textrm{for all $\beta$}
+\end{equation*}
+and as a result $f(x)$ must itself be quite small,
+\begin{align*}
+\bigl|f(x)\bigr| & = \biggl|\sum_{\abs{\alpha} = 0}^{s - 1} f^{(\alpha)}(x_0) (x - x_0)^{\alpha} + \sum_{\abs{\alpha} = s}(x - x_0)^{\alpha}F_{\alpha}(x_0,x)\biggr| \\
+& = \biggl|\sum_{\abs{\alpha} = s}(x - x_0)^{\alpha}F_{\alpha}(x_0,x)\biggr| \\
+& \leq c r^{s} \abs{F_{\alpha}(x_0,x)} \leq c r^{s} \norm{f}_{C^s(\Xset)}.
+\end{align*}
+Observe $\Ebb\Bigl[D_kf(x)\Bigr] = 0$ unless $x_j \in B(x,qr)$ for each $j \in k$, which implies $x_j \in \Xset \setminus \Xset_{2qr}$. Thus,
+\begin{equation*}
+\abs{\Ebb\Bigl[D_kf(x)\Bigr]} \leq 2^q p_{\max}^q \max_{\wt{x} \in x_k, x} \bigl\{\abs{f(\wt{x})}\bigr\}  \leq c r^s \norm{f}_{C^s(\Xset)},
+\end{equation*}
+establishing~\eqref{eqn:leading_term_holder_compact_2} for $x \in \Xset \setminus \Xset_{qr}$.
+
+\subsubsection{Interior point.}
+To show the desired result when $x \in \Xset_{qr}$ and $s \geq 1$, we use induction on $q$.
+
+\paragraph{Base case.}
+In the base case $q = 1$, meaning $D_kf$ is only a single-difference operator.
+Since $s \geq 1$, replacing $f$ by its Taylor expansion inside the first order expected difference operator $\Ebb[D_kf(x)]$ yields
+\begin{equation}
+\label{eqn:leading_term_holder_compact_pf1}
+\Ebb\Bigl[D_kf(x)\Bigr] = \sum_{1 \leq \abs{\alpha} < s} \Ebb_{\alpha,x} \cdot f^{(\alpha)}(x)  + \sum_{\abs{\alpha} = s} \Bigl(\Ebb_{\alpha}\bigl[F_{\alpha}\bigr]\Bigr)(x)
+\end{equation}
+When $s = 1$ only the second term in the previous expression is non-zero, and we therefore begin by analyzing this term, obtaining that for any $x \in \Xset$ and each $\abs{\alpha} = s$,
+\begin{align}
+\biggl|\Bigl(\Ebb_{\alpha}\bigl[F_{\alpha}\bigr]\Bigr)(x)\biggr| & = \biggl|\int (y - x)^{\alpha} F_{\alpha}(x,y) K_r(y,x) p(y) \,dy\biggr| \nonumber \\
+& \leq r^s p_{\max} \abs{F_{\alpha}(x,y)} \nonumber \\
+& \leq r^s p_{\max} \norm{f}_{C^s(\Xset)}. \label{eqn:leading_term_holder_compact_pf2}
+\end{align}
+Therefore $\Ebb_{\alpha}\bigl[F_{\alpha}\bigr] \in C^0(\Xset)$ additionally satisfies~\eqref{eqn:leading_term_holder_compact_2}, and the claim of Lemma~\ref{lem:leading_term_holder_compact} is established for $q = 1$ and $s = 1$.
+
+When $s \geq 2$ we must analyze $\Ebb_{\alpha,x} \cdot f^{(\alpha)}(x)$, which we do by using the Taylor expansion of $p$.
+Since $B(x,r) \subset \Xset$, we recall that $\Ibb_{\alpha + \beta,x} = r^{\abs{\alpha} + \abs{\beta}}I_{\alpha,\beta}$; thus
+\begin{align*}
+\Ebb_{\alpha,x} & = \int(y - x)^{\alpha} K_r(y,x) p(y) \,dy \\
+& = \sum_{\abs{\beta} = 0}^{s - 2} p^{(\beta)}(x) \Ibb_{\alpha + \beta,x} + \sum_{\abs{\beta} = s - 1} \Bigl(\Ibb_{\alpha + \beta}\bigl[G_{\beta}\bigr]\Bigr)(x) \\
+& = \sum_{\abs{\beta} = 0}^{s - 2} p^{(\beta)}(x) r^{\abs{\alpha} + \abs{\beta}} I_{\alpha + \beta} + \sum_{\abs{\beta} = s - 1} \Bigl(\Ibb_{\alpha + \beta}\bigl[G_{\beta}\bigr]\Bigr)(x).
+\end{align*}
+Replacing $\Ebb_{\alpha,x}$ by this expansion in~\eqref{eqn:leading_term_holder_compact_pf1} gives
+\begin{equation*}
+\Ebb\Bigl[D_kf(x)\Bigr] = \sum_{\abs{\alpha} = 1}^{s - 1} \sum_{\abs{\beta} = 0}^{s - 2} r^{\abs{\alpha} + \abs{\beta}} I_{\alpha + \beta} f^{(\alpha)}(x) p^{(\beta)}(x)  + \sum_{\abs{\alpha} = 1}^{s} \sum_{\abs{\beta} = s - 1} \Bigl(\Ibb_{\alpha + \beta}\bigl[G_{\beta}\bigr]\Bigr)(x) f^{(\alpha)}(x) + \sum_{\abs{\alpha} = s} \Bigl(\Ebb_{\alpha}\bigl[F_{\alpha}\bigr]\Bigr)(x)
+\end{equation*}
+We now divide the sum in the first term based on the size of $\abs{\alpha} + \abs{\beta}$.
+The critical fact is that $I_{\alpha + \beta} = 0$ when $\abs{\alpha} + \abs{\beta} = 1$.
+When $s = 2$ this leaves
+\begin{equation*}
+\Ebb\Bigl[D_kf(x)\Bigr] = \sum_{\abs{\alpha} = 1}^{s} \sum_{\abs{\beta} = s - 1} \Bigl(\Ibb_{\alpha + \beta}\bigl[G_{\beta}\bigr]\Bigr)(x) f^{(\alpha)}(x) + \sum_{\abs{\alpha} = s} \Bigl(\Ebb_{\alpha}\bigl[F_{\alpha}\bigr]\Bigr)(x).
+\end{equation*}
+In~\eqref{eqn:leading_term_holder_compact_pf2} we have already shown that the second term belongs to $C^{0}(\Xset)$, and provided an appropriate upper bound on its norm.
+Similar analysis shows that when $1 \leq \abs{\alpha} \leq s$ and $\abs{\alpha} + \abs{\beta} \geq s$,
+\begin{align}
+\norm{\Ibb_{\alpha + \beta}\bigl[G_{\beta}\bigr]f^{(\alpha)}}_{C^0(\Xset)} & \leq r^{\abs{\alpha} + \abs{\beta}} p_{\max} \norm{f^{(\alpha)}}_{C^0(\Xset)} \nonumber \\
+& \leq r^{\abs{\alpha} + \abs{\beta}} p_{\max} \norm{f}_{C^s(\Xset)} \nonumber \\
+& \leq r^s p_{\max} \norm{f}_{C^s(\Xset)}, \label{eqn:leading_term_holder_compact_pf5}
+\end{align}
+taking care of the first term, and establishing~\eqref{eqn:leading_term_holder_compact_2} when $s = 2$. Otherwise when $s > 2$, we rearrange
+\begin{equation*}
+\sum_{\abs{\alpha} = 1}^{s} \sum_{\abs{\beta} = 0}^{s - 2} r^{\abs{\alpha} + \abs{\beta}} I_{\alpha + \beta} f^{(\alpha)}(x) p^{(\beta)}(x) = \sum_{\ell = 2}^{s - 1} r^{\ell} \Biggl\{\underbrace{\sum_{\abs{\alpha} + \abs{\beta} = \ell} I_{\alpha + \beta} f^{(\alpha)}(x) p^{(\beta)}(x)}_{:=f_{\ell,1}(x)}\Biggr\} + \sum_{\ell = s + 1}^{2s - 2} r^{\ell} \sum_{\abs{\alpha} + \abs{\beta} = \ell} I_{\alpha + \beta} f^{(\alpha)}(x) p^{(\beta)}(x).
+\end{equation*}
+and therefore
+\begin{align*}
+& \abs{\Ebb\Bigl[D_kf(x)\Bigr] - \sum_{\ell = 2}^{s - 1} r^{\ell} f_{\ell,1}(x)} = \\ & ~~~~ \sum_{\ell = s + 1}^{2s - 2} r^{\ell} \sum_{\abs{\alpha} + \abs{\beta} = \ell} I_{\alpha + \beta} f^{(\alpha)}(x) p^{(\beta)}(x) +  \sum_{\abs{\alpha} = 1}^{s} \sum_{\abs{\beta} = s - 1} \Bigl(\Ibb_{\alpha + \beta}\bigl[G_{\beta}\bigr]\Bigr)(x) f^{(\alpha)}(x) + \sum_{\abs{\alpha} = s} \Bigl(\Ebb_{\alpha}\bigl[F_{\alpha}\bigr]\Bigr)(x)
+\end{align*}
+On the left hand side, taking $\ell = \abs{\alpha} + \abs{\beta}$, note that for each $\ell < s$ the function $f_{\ell,1} \in C_c^{s - \ell}(\Xset)$ and further
+\begin{equation}
+\label{eqn:leading_term_holder_compact_pf3}
+\norm{f_{\ell,1}}_{C^{s - \ell}(\Xset)} \leq c p_{\max} \norm{f}_{C^s(\Xset)}.
+\end{equation}
+
+The right hand side consists of three terms, and we have already obtained sufficient estimates on the second and third term in~\eqref{eqn:leading_term_holder_compact_pf2} and~\eqref{eqn:leading_term_holder_compact_pf5}, respectively.
+It remains to deal with the first term.
+We have that $f^{(\alpha)}\cdot p^{(\beta)} \in C_c^{0}(\Xset)$ and
+\begin{equation}
+\label{eqn:leading_term_holder_compact_pf4}
+\norm{f^{(\alpha)}p^{(\beta)}}_{C^{0}(\Xset)} \leq p_{\max} \norm{f}_{C^s(\Xset)},
+\end{equation}
+establishing~\eqref{eqn:leading_term_holder_compact_1} when $s > 3$.
+
+\paragraph{Induction Step.}
+
+We now assume that~\eqref{eqn:leading_term_holder_compact_1} and~\eqref{eqn:leading_term_holder_compact_2} hold for all $k \in (n)^q$ and $x \in \Xset_{qr}$, and prove the desired estimates hold with respect to $\Ebb[D_jD_kf(x)]$ for all $(kj) \in (n)^{q + 1}$ and $x \in \Xset_{(q+1)r}$.
+
+We first consider the case where $s \leq 2q$. Here,
+\begin{equation*}
+\abs{\Ebb\Bigl[D_jD_kf(x)\Bigr]} \leq 2p_{\max} \sup_{x \in \Xset} \biggl\{\Ebb[D_kf(x)]\biggr\} \leq c r^s \norm{f}_{C^s(\Xset)}
+\end{equation*}
+--with the last inequality following by the inductive hypothesis--and~\eqref{eqn:leading_term_holder_compact_2} is established.
+
+Otherwise $s \geq 2q + 1$.
+Our strategy will be to apply the inductive hypothesis twice, once to expand the $q$th order operator $D_k$ and once to expand the first order operator $D_j$.
+To achieve the former, we first note that by the law of iterated expectation
+\begin{equation}
+\label{eqn:leading_term_holder_compact_pf6}
+\Ebb\Bigl[D_jD_kf(x)\Bigr] = \Ebb\biggl[\Bigl(\Ebb\bigl[D_kf(x_j)|x_j\bigr] - \Ebb\bigl[D_kf(x)\bigr]\Bigr)K_r(x_j,x)\biggr].
+\end{equation}
+We now apply the inductive hypothesis to both $\Ebb\Bigl[D_kf(x)\Bigr]$ and  $\Ebb\Bigl[D_kf(x_j)|x_j\Bigr]$, which states that
+\begin{equation*}
+\abs{\Ebb\Bigl[D_kf(x)\Bigr] - \sum_{\ell = 2q}^{s - 1} r^{\ell}f_{\ell,q}(x)}, \abs{K_r(x_j,x)\biggl(\Ebb\Bigl[D_kf(x_j)|x_j\Bigr] - \sum_{\ell = 2q}^{s - 1} r^{\ell}f_{\ell,q}(x_j)\biggr)} \leq cr^s \norm{f}_{C^s(\Xset)}.
+\end{equation*}
+We emphasize that we may imply the inductive hypothesis in the latter case since $x \in \Xset_{(q+1)r}$ implies either $K_r(x_j,x) = 0$, or $x_j \in \Xset_{qr}$.
+Plugging back in to~\eqref{eqn:leading_term_holder_compact_pf6} gives
+\begin{equation}
+\label{eqn:leading_term_holder_compact_pf7}
+\abs{\Ebb\Bigl[D_jD_kf(x)\Bigr] - \sum_{\ell = 2q}^{s - 1} r^{\ell} \Ebb\Bigl[D_jf_{\ell,q}(x)\Bigr]} \leq c r^s \norm{f}_{C^s(\Xset)}.
+\end{equation}
+Next, we note that $f_{\ell,q} \in C_c^{s-\ell}(\Xset)$, for each $\ell = 2q,\ldots,s - 1$. We now apply the inductive hypothesis again, this time to $\Ebb[D_jf_{\ell,q}(x)]$. If $s - \ell \leq 2$,
+\begin{equation*}
+\abs{\Ebb\Bigl[D_jf_{\ell,q}(x)\Bigr]} \leq c r^{s - \ell} \norm{f_{\ell,q}}_{C^{s - \ell}(\Xset)} \leq c r^{s - \ell} \norm{f}_{C^{s}(\Xset)}.
+\end{equation*}
+Therefore when $s = 2q + 1$ or $s = 2q + 2$, by~\eqref{eqn:leading_term_holder_compact_pf7}
+\begin{equation*}
+\abs{\Ebb\Bigl[D_jD_kf(x)\Bigr]} \leq \sum_{\ell = 2q}^{s - 1} r^{\ell}\abs{\Ebb\Bigl[D_jf_{\ell,q}(x)\Bigr]} + c r^s \norm{f}_{C^s(\Xset)} \leq cr^s\norm{f}_{C^s(\Xset)}
+\end{equation*}
+completing the proof of~\eqref{eqn:leading_term_holder_compact_2}.
+
+On the other hand if $s - \ell > 2$, there exist functions $\wt{f}_{\ell + \wt{\ell},q} \in C_c^{s - \ell - \wt{\ell}}(\Xset)$ satisfying $\norm{\wt{f}_{\ell + \wt{\ell},q}}_{C^{s - \ell}(\Xset)} \leq c \norm{f_{\ell,q}}_{C^s(\Xset)}$ such that
+\begin{equation*}
+\abs{\Ebb\Bigl[D_jf_{\ell,q}(x)\Bigr] - \sum_{\ell = 2}^{s - \ell  - 1} r^{\wt{\ell}} f_{\ell + \wt{\ell},q}(x)} \leq c r^{(s - \ell - \wt{\ell})} \norm{f_{\ell,q}}_{C^{s - \ell}(\Xset)} \leq c r^{(s - \ell - \wt{\ell})} \norm{f}_{C^{s}(\Xset)}.
+\end{equation*}
+Plugging in to~\eqref{eqn:leading_term_holder_compact_pf7}, we have
+\begin{equation*}
+\abs{\Ebb\Bigl[D_jD_kf(x) - \sum_{\ell = 2q}^{s - 3} \sum_{\wt{\ell} = 2}^{s - \ell - 1} \wt{f}_{\ell + \wt{\ell},q}(x) r^{\ell + \wt{\ell}}\Bigr]} \leq cr^s \norm{f}_{C^s(\Xset)}
+\end{equation*}
+and rewriting the sum inside the absolute value over $\ell + \wt{\ell} = 2q + 2,\ldots,s - 1$, we have completed the proof of~\eqref{eqn:leading_term_holder_compact_1}.
+
+\section{Additional Results}
+
+To show equation~\eqref{eqn:leading_term_sobolev_compact_1} in Lemma~\ref{lem:leading_term_sobolev_compact}, it suffices to prove an equivalent bound on the $\Leb^2$ norm $\norm{f}_{\Leb^2(U_{2qr})}$.
+We show this in the case where the domain is the positive halfspace in $\Rd$, that is $\Xset = R_+^d = \set{x \in \Reals^d: x_d > 0}$.
+
+\begin{lemma}
+	Suppose $g \in C_c^{s}(\Rdp) \cap H^s(\Rdp)$. Then
+	\begin{equation}
+	\norm{g}_{L^2(U_{r})} \leq 2 r^s\norm{g}_{H^s(\Rdp)}
+	\end{equation}  
+\end{lemma}
+\begin{proof}
+	We immediately take advantage of the axis-oriented structure of $\Rdp$.
+Write $x \in U_{2qr}$ as $x = (x',te_d)$, where $x' \in \Reals^{d - 1} \cap \{x_d = 0\}$ lies on the plane at $x_d = 0$, and $0 \leq \abs{t} \leq r$.
+Since $g$ is compactly supported on $\Rdp$, by taking a Taylor expansion we obtain
+	\begin{align*}
+	g(x) & = \sum_{\ell = 0}^{s - 1} \frac{\partial^\ell}{\partial e_d^\ell}g(x') t^{\ell} + \int_{0}^{t} \frac{\partial^s}{\partial e_d^s} g(x' + he_d) t^{s - 1} \,dh \\
+	& = \int_{0}^{t} \frac{\partial^s}{\partial e_d^s} g(x' + he_d) h^{s - 1} \,dh
+	\end{align*}
+	We then bound the squared $\Leb^2$ norm of $g$ using Hardy's inequality,
+	\begin{align*}
+	\int_{\Xset} \bigl(g(x)\bigr)^2 \,dx & = \int_{\Reals^{d - 1}} \int_{0}^{r} \biggl(\int_{0}^{t} \frac{\partial^s}{\partial e_d^s} g(x' + he_d) h^{s - 1} \,dh\biggr)^2 \,dt \,dx' \\
+	& \leq r^{2s} \int_{\Reals^{d - 1}} \int_{0}^{r} \biggl(\frac{1}{t}\int_{0}^{t} \frac{\partial^s}{\partial e_d^s} g(x' + he_d) \,dh\biggr)^2 \,dt \,dx' \\
+	& \leq 4 r^{2s} \int_{\Reals^{d - 1}} \int_{0}^{\infty} \Bigl(\frac{\partial^s}{\partial e_d^s} g(x' + te_d)\Bigr)^2  \,dt \,dx' \\
+	& \leq 4 r^{2s} \norm{g}_{H^s(\Rdp)}^2.
+	\end{align*}
 \end{proof}
 
-Here are a couple of Lemmas which helped us with the proof of Lemma~\ref{lem:leading_term_sobolev_compact}.
+Now we show a similar result when $\Xset \subset \Rd$ is assumed to be a bounded open set with Lipschitz boundary. The proof is slightly more complicated, but the key ideas are the same.
 
@@
 
-\subsection{One-Sided Concentration}
-
-The proof of Lemma~\ref{lem:empirical_norm_sobolev} relies on (a variant of) the Paley-Zygmund Inequality.
-\begin{lemma}
-	\label{lem:paley_zygmund}
-	Let $f$ satisfy the following moment inequality for some $b \geq 1$:
-	\begin{equation}
-	\label{eqn:paley_zygmund_1}
-	\Ebb\bigl[\norm{f}_n^4\bigr] \leq \left(1 + \frac{1}{b^2}\right)\cdot\Bigl(\Ebb\bigl[\norm{f}_n^2\bigr]\Bigr)^2.
-	\end{equation}
-	Then,
-	\begin{equation}
-	\label{eqn:paley_zygmund_2}
-	\mathbb{P}\left[\norm{f}_n^2 \geq \frac{1}{b} \Ebb\bigl[\norm{f}_n^2\bigr]\right] \geq 1 - \frac{5}{b}.
-	\end{equation}
-\end{lemma}
-\begin{proof}
-	Let $Z$ be a non-negative random variable such that $\mathbb{E}(Z^q) < \infty$. The Paley-Zygmund inequality says that for all $0 \leq \lambda \leq 1$,
-	\begin{equation}
-	\label{eqn:paley_zygmund_pf1}
-	\mathbb{P}(Z > \lambda \mathbb{E}(Z^p)) \geq \left[(1 - \lambda^p) \frac{\mathbb{E}(Z^p)}{(\mathbb{E}(Z^q))^{p/q}}\right]^{\frac{q}{q - p}}
-	\end{equation}
-	Applying~\eqref{eqn:paley_zygmund_pf1} with $Z = \norm{f}_n^2$, $p = 1$, $q = 2$ and $\lambda = \frac{1}{b}$, by assumption~\eqref{eqn:paley_zygmund_1} we have
-	\begin{equation*}
-	\mathbb{P}\Bigl(\norm{f}_n^2 > \frac{1}{b} \mathbb{E}[\norm{f}_n^2]\Bigr) \geq \Bigl(1 - \frac{1}{b}\Bigr)^2 \cdot  \frac{\bigl(\mathbb{E}[\norm{f}_n^2]\bigr)^2}{\mathbb{E}[\norm{f}_n^4]} \geq \frac{\Bigl(1 - \frac{2}{b}\Bigr)}{\Bigl(1 + \frac{1}{b^2}\Bigr)} \geq 1 - \frac{5}{b}.
-	\end{equation*}
-\end{proof}
-
 \end{document}

### 2020-02-24 notes/graph_testing.tex
updated main notes

@@
 	\label{event:discrete_sobolev_norm}
-	\textbf{Graph Sobolev norm:} For any $n^{-1/(2(s - 1) + d)}\leq r$ such that $r \to 0$ as $n \to \infty$,
+	\textbf{Graph Sobolev norm:} There exists an $r^{\star}$ (which does not depend on $f$) such that for any $n^{-1/(2(s - 1) + d)}\leq r \leq r^{\star}$,
 	\begin{equation}
@@
 	\end{equation}
-	for all $n$ sufficiently large.
 	\item 
@@
 	\label{lem:remainder_term_sobolev}
-	Suppose $f \in \Leb^2(U_r)$ and $k \in (n)$. Then,
+	Let $U \subset \Rd$ be an open set, and $U_r = \set{x \in \Rd: \mathrm{dist}(x,U) < x}$. Suppose $f \in \Leb^2(U_r)$ and $k \in (n)$. Then,
 	\begin{equation*}

### 2020-03-05 notes/weekly_notes/3.5.20.NotesForWeek.tex
added notes for this week

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\makeatletter
\newcommand{\vast}{\bBigg@{4}}
\newcommand{\Vast}{\bBigg@{5}}
\makeatother

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{\mathcal{L}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 2/28/20 - 3/5/20}
\author{Alden Green}
\date{\today}
\maketitle

Let $N$ be an integer, $n = N^d$, and 
\begin{equation*}
\wb{X} := \Bigl\{\frac{1}{N}(j_1,\ldots,j_d): j \in [N]^d\Bigr\}
\end{equation*}
consist of $n$ total evenly spaced grid points on $[0,1]^d$. We observe $n$ samples according to regression model
\begin{equation}
\label{eqn:grid_regression_model}
y_j = f(\wb{x}_j) + \varepsilon_j, ~\varepsilon_j \overset{\mathrm{i.i.d}}{\sim} \mathcal{N}(0,1) ~~\textrm{for each $j \in [N]^d$,}
\end{equation}
and our goal is to perform the goodness-of-fit testing
\begin{equation*}
\mathbb{H}_0 : f = f_0 := 0 ~~\textrm{vs.}~~ \mathbb{H}_a: f \neq f_0.
\end{equation*}

For an arbitrary graph $G$ on $n$ nodes with graph Laplacian $L_G = D_G - A_G$, let $\Lambda(G)$ consist of the eigenvalues $0 = \lambda_1(G) \leq \lambda_2(G) \leq \cdots \leq \lambda_n(G)$ of $L_G$, with $v_k(G) = (v_{k,1}(G),\ldots,v_{k,n}(G)) \in \Reals^n$ denoting the eigenvector corresponding to the $i$th eigenvalues $\lambda_k(G)$. The graph Laplacian eigenvector projection test is a linear projection test; letting $\kappa$ be some integer between $1$ and $n$, we define $\Pi_{k,G}(\theta) := \frac{1}{\sqrt{n}}\sum_{k = 1}^{\kappa} \dotp{v_k}{\theta}$ to be a suitable rescaling of the projection of $\theta$ onto the span of $v_1(G),\ldots,v_k(G)$.

Our test statistic will be
\begin{equation*}
T_{\mathrm{spec}}(G) := \norm{\Pi_{\kappa,G}(Y)}_n^2 - \frac{\kappa}{n},
\end{equation*}
a simple function of the projection of the data $Y$ onto the span of the first $\kappa$ eigenvectors of the graph Laplacian $L(G)$; we will suitably choose $G$ based on the assumptions we make on $f$.

\section{Sobolev testing with fixed design.}
For now, we confine our attention to the periodic Sobolev space $H_{\textrm{per}}^{s}([0,1]^d)$ which consists of all those functions $f \in \Leb^2([0,1]^d)$ whose tensor product Fourier coefficients decay at a sufficiently fast rate. When $d = 1$ and $x \in [0,1]$, let
\begin{align*}
\phi_1(x) & = 1 \\
\phi_{2k}(x) & = \cos(2k\pi x),~~ k \in \mathbb{N} \\
\phi_{2k + 1}(x) & = \sin(2k\pi x),~~ k \in \mathbb{N}
\end{align*}
be the trigonometric basis over $[0,1]$. The tensor product Fourier basis is defined on $[0,1]^d$ as
\begin{equation*}
\phi_k(x) = \prod_{i = 1}^{d} \phi_{k_i}(x_i),
\end{equation*}
the Sobolev space $H_{\textrm{per}}^{s}([0,1]^d)$ as 
\begin{equation*}
H^{\textrm{per}}([0,1]^d) = \set{\sum_{k \in \mathbb{N}^d} \theta_k \phi_k^{(d)}: \sum_{k \in \mathbb{N}^d} \theta_k^2 \abs{k}^{2s} < \infty},
\end{equation*}
and the unit ball $H_{\textrm{per}}^{s}([0,1]^d, L)$ as 
\begin{equation*}
H^{\textrm{per}}([0,1]^d;L) = \set{\sum_{k \in \mathbb{N}^d} \theta_k \phi_k^{(d)}: \sum_{k \in \mathbb{N}^d} \theta_k^2 \abs{k}^{2s} \leq L},
\end{equation*}

Intuitively, a good choice of graph $G$ for our test statistic $T_{\spec}(G)$ will be one which keeps the approximation error
\begin{equation*}
\norm{f}_{\Leb^2}^2 - \Ebb\Bigl[\norm{\Pi_{\kappa,G}(Y)}_{n}^2\Bigr] =  \norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,G}(f)}_{n}^2
\end{equation*}
small when $\kappa$ is not too large.

\subsection{Torus Graph}

We will see that the torus graph $\wt{G}_d$ fulfills this purpose. The $1$-dimensional torus (ring) graph is defined as
\begin{equation*}
\wt{G}_1 = (\wb{X}; \wt{E}),~~ \wt{E} = \Bigl\{(\wb{x}_i,\wb{x}_j): i \in [n], j = (i + 1) \mod n\Bigr\};
\end{equation*}
when $d > 1$, we define $\wt{G}_d = \wt{G}_1 \otimes \cdot \wt{G}_1$ as the tensor product of $\wt{G}_1$ iterated $d$ times.

\begin{lemma}
	\label{lem:grid_sobolev_approximation_error}
	Let $K := \kappa^{1/d}$. For any $2s > d$ and $f \in H_{\textrm{per}}^s([0,1]^d;L)$, 
	\begin{equation}
	\label{eqn:grid_sobolev_approximation_error}
	\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,G}(f)}_{n}^2 \leq c n^{1/2 - s/d} \kappa^{1/2} \norm{f}_2 + \sum_{k \in \mathbb{N}^d - [K]^d} \theta_k^2
	\end{equation}
	for a constant $c$ which depends only on $L$ and $d$. 
\end{lemma}

It is possible to improve on Lemma~\ref{lem:grid_sobolev_approximation_error} under \textcolor{red}{(relatively mild)} additional assumptions. Specifically, we assume the existence of $\varepsilon_j$ satisfying $j \varepsilon_j$ is monotone non-increasing in $j$, such that for every $f \in H_{\textrm{per}}^{s}([0,1]^d)$, $f = \sum_{k \in \mathbb{N}^d} \theta_k \phi_k$,
\begin{equation}
\label{asmp:fourier_coefficient_decay}
\abs{\theta_k} \leq \prod_{i = 1}^{d} \varepsilon_{k_i}~~\textrm{for all $k \in \mathbb{N}^d$},~~ \sum_{j = 1}^{\infty} \varepsilon_j = C,~~ \sum_{j = 1}^{\infty} \varepsilon_j^2 \leq C \biggl(\sum_{k \in \mathbb{N}^d} \theta_k^2\biggr)^{1/d}
\end{equation}
for some constant $C < \infty$.

\textcolor{red}{(TODO:)} {Unfortunately these additional assumptions are not mild at all, but instead relatively close to assuming the function is $s$-Holder.}

\begin{lemma}
	\label{lem:grid_sobolev_approximation_error_2}
	Suppose that~\eqref{asmp:fourier_coefficient_decay} holds with respect to $H_{\textrm{per}}^{s}([0,1]^d)$. Then for every $f \in H_{\textrm{per}}^{s}([0,1]^d)$,
	\begin{equation}
	\label{eqn:grid_sobolev_approximation_error_2}
	\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,G}(f)}_{n}^2 \leq c \norm{f}_{\Leb^2}^2 \Biggl( \frac{K^{1/2}}{N \norm{f}_{\Leb^2}^{1/d}} + \frac{K^{d/2}}{n\norm{f}_{\Leb^2}}\Biggr) + \sum_{k \in \mathbb{N}^d - [K]^d} \theta_k^2
	\end{equation}
\end{lemma}
When we choose $\kappa = n^{2d/(4s + d)}$, and additionally $\norm{f}_{\Leb^2} \geq n^{-2s/(4s + d)}$, we note that 
\begin{equation*}
\frac{K^{1/2}}{N \norm{f}_{\Leb^2}^{1/d}} + \frac{K^{d/2}}{n\norm{f}_{\Leb^2}} = o(1)
\end{equation*}
and therefore the discretization error is asymptotically negligible.

Now we will attempt to find a sharper bound on discretization error than Lemma~\ref{lem:grid_sobolev_approximation_error}, but that requires less severe assumptions than Lemma~\ref{lem:grid_sobolev_approximation_error_2}.
\begin{lemma}
	\label{lem:grid_sobolev_approximation_error_3}
	For any $2s > d$ and $f \in H_{\textrm{per}}^s([0,1]^d;L)$, 
	\begin{equation*}
	\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,G}(f)}_{n}^2 \leq c n^{1/2 - s/d} \norm{f}_2 + \sum_{k \in \mathbb{N}^d - [K]^d} \theta_k^2
	\end{equation*}
\end{lemma}

\subsection{Grid graph}
In this section we will find it convenient to assume the grid points are symmetric in $[0,1]^d$, so we now treat
\begin{equation*}
\wb{X} := \set{\frac{1}{2N}(2j_1 - 1,\ldots,2j_d - 1): j \in [N]^d}.
\end{equation*}

The grid graph $\wb{G}_d = \wb{G} \otimes \cdots \otimes \wb{G}$, where
\begin{equation*}
\wb{G} = (\wb{X}; \wb{E}),~~ \wb{E} = \Bigl\{(\wb{x}_i,\wb{x}_j): i \in [n - 1], j = (i + 1)\Bigr\}; 
\end{equation*}
is the path (i.e. the torus with the edge connecting $1$ to $n$ removed.) The eigenvectors of the path can be written as follows:
\begin{equation*}
v_{k,i}(\wb{G}) = \cos\Bigl(\frac{\pi (k - 1) (i - 1/2)}{n}\Bigr), \textrm{for $i = 1,\ldots,n$ and $k = 1,\ldots,n - 1$}
\end{equation*}
with the usual consequence for the eigenvectors of the grid:
\begin{equation*}
v_{k,i}(\wb{G}_d) = \prod_{j = 1}^d v_{k_j,i_j}(\wb{G}),~~ \textrm{for $i \in [N]^d$, $j \in [N]^d$.}
\end{equation*}
We extend these eigenvectors to an orthonormal system over $\Leb^2([0,1]^d)$ as follows. Let
\begin{equation*}
\varphi_k(x) = \sqrt{2}\cos\biggl(\pi (k - 1) x\biggr),~~\textrm{for $k \in \Nbb, x \in [0,1]$},~~\textrm{and}~~ \varphi_k(x) = \prod_{j = 1}^{d} \varphi_{k_j}(x_j),~~\textrm{for $k \in \Nbb^d, x \in [0,1]^d$},
\end{equation*}
and note that by shifting the location of the design points $\wb{X}$, we have that $\varphi_k(\wb{x}_i) = \sqrt{n} v_{k,i}(\wb{G}_d)$ for all $k,i \in [N]^d$.

We let the space 
\begin{equation*}
\wt{H}^{s}([0,1]^d;L) = \Bigl\{ \sum_{k \in \mathbb{N}^d} \theta_k \varphi_k: \sum_{k \in \mathbb{N}^d} \theta_k^2 a_k^{2} < L \Bigr\}
\end{equation*}
where in the one-dimensional case,
\begin{equation*}
a_{k - 1} =
\begin{cases*}
(k - 1)^s,~~ \textrm{if $k - 1$ is even} \\
(k - 2)^s,~~ \textrm{if $k - 1$ is odd}
\end{cases*}
\end{equation*}
and for general $d$, $a_k = \sqrt{\sum_{j = 1}^{d} a_{k_j}^2}$. It is not hard to show that $\wt{H}^s([0,1]^d)$ consists of those functions in $H^s([0,1]^d)$ which satisfy some boundary conditions
\begin{equation*}
\wt{H}^s([0,1]^d;L) = \Bigl\{f \in H^{s}([0,1]^d;L'): \forall x_0 \in \partial [0,1]^d, \frac{\partial^{\ell} f}{\partial \mathbf{n}^{\ell}}(x_0) = 0,~~\textrm{for $\ell = 1,\ldots,s-1$} \Bigr \} \supseteq H_0^s([0,1]^d;L')
\end{equation*}
for a constant $L'$ which depends only on $L$ and $s'$; we work through the details for the case $d = 1$ in Section~\ref{subsec:sobolev_class_equivalence}.

For functions $f \in \wt{H}^{s}([0,1]^d;L)$, the error in approximating the continuum norm $\norm{f}_{\Leb^2}$ by the discrete norm of a projection of $f$ onto the eigenvectors of the grid is not too large.

\begin{lemma}
	\label{lem:grid_sobolev_approximation_error_4}
	Suppose $2s < d$ and $f = \sum_{k \in \mathbb{N}^k} \theta_k \varphi_k$ satisfies $f \in \wt{H}_{\textrm{per}}^{s}([0,1]^d;L)$. Then
	\begin{equation*}
	\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,\wb{G}_d}(f)}_{n}^2 \leq c n^{1/2 - s/d} \norm{f}_2 + \sum_{k \in \mathbb{N}^d - [K]^d} \theta_k^2
	\end{equation*}
\end{lemma}

\section{Proofs}
\subsection{Proof of Lemma~\ref{lem:grid_sobolev_approximation_error}}
Recall that the eigenvectors of the torus graph $v_k := v_k(\wt{G}_d)$ satisfy
\begin{equation*}
v_{k}(\wb{x}_i) = \frac{1}{\sqrt{n}} \phi_{k}(\wb{x}_i)
\end{equation*}
for each $i \in [N]^d$ and $k \in \mathbb{N}^d$. Rewriting $\Pi_{\kappa,G}$ as a function of the tensor product Fourier basis, we obtain
\begin{align*}
\Pi_{\kappa,G}(f) & = \frac{1}{n} \sum_{k \in [K]^d} \Bigl\{\sum_{i \in [N]^d} f(\wb{x}_i) \phi_k(\wb{x}_i)\Bigr\} \phi_k \\
& :=  \sum_{k \in [K]^d} \wt{\theta}_k \phi_k
\end{align*}
where the right hand side is interpreted as a function over $\wb{X}$. Since $\phi_k$ are $L_2(\wb{X})$ orthonormal, we obtain
\begin{align*}
\norm{\Pi_{\kappa,G}(f)}_n^2 & = \sum_{k \in [K]^d} \wt{\theta}_k^2 \\ 
& = \sum_{k \in [K]^d} {\theta}_k^2 + \sum_{k \in [K]^d} \wt{\theta}_k^2 - \theta_k^2 \\
& = \norm{f}_{\Leb^2([0,1]^d)}^2 - \sum_{k \in \mathbb{N}^d - [K]^d} \theta_k^2 + \sum_{k \in [K]^d} \wt{\theta}_k^2 - \theta_k^2
\end{align*}
and rearranging, we have
\begin{equation*}
\norm{f}_{\Leb^2([0,1]^d)}^2 - \norm{\Pi_{\kappa,G}(f)}_n^2 = \sum_{k \in [K]^d} \theta_k^2 - \wt{\theta}_k^2 + \sum_{k \in \mathbb{N}^d - [K]^d} \theta_k^2
\end{equation*}
The second term in the above equation corresponds to the second term in~\eqref{eqn:grid_sobolev_approximation_error}, and so we focus our attention on the first term, which represents discretization error. Applying the parallelogram law and the Cauchy-Schwarz inequality gives
\begin{equation}
\label{eqn:grid_sobolev_approximation_error_pf2}
\begin{aligned}
\sum_{k \in [K]^d} \theta_k^2 - \wt{\theta}_k^2 & \leq \sum_{k \in [K]^d} \max\Bigl\{(\theta_k + \wt{\theta}_k)(\theta_k - \wt{\theta}_k),0\Bigr\} \\
& \leq \left(\sum_{k \in [K]^d} \max\Bigl\{(\theta_k + \wt{\theta}_k)^2,0\Bigr\}\right)^{1/2} \left(\sum_{k \in [K]^d} (\theta_k - \wt{\theta}_k)^2\right)^{1/2} \\
& \leq \sqrt{2} \norm{f}_{\Leb^2([0,1]^d)} \left(\sum_{k \in [K]^d} (\theta_k - \wt{\theta}_k)^2\right)^{1/2}
\end{aligned}
\end{equation}

We focus on the differences $\theta_k - \wt{\theta}_k$, and using the same steps as \textcolor{red}{(Tsybakov)} Lemma 1.8 we obtain
\begin{equation}
\label{eqn:grid_sobolev_approximation_error_pf1}
\max_{k \in [K]^d}\abs{\theta_k - \wt{\theta}_k} \leq 2 L^{1/2} \left(\sum_{j \in \mathbb{N}^d - [N]^d} \abs{j}^{-2s}\right)^{1/2}
\end{equation}
so long as $K^d \leq n - 1$. Upper bounding the Riemann sum in~\eqref{eqn:grid_sobolev_approximation_error_pf1} by a corresponding integral, and converting to spherical coordinates, we have
\begin{align*}
\sum_{j \in \mathbb{N}^d - [N]^d} \abs{j}^{-2s} & \leq d \sum_{j \in \mathbb{N}^d: j_1 \geq N + 1} \abs{j}^{-2s} \\
& \leq d \int_{\Reals^d \setminus B(0,N+1)} \abs{x}^{-2s} \,dx \\
& \leq c \int_{N + 1}^{\infty} r^{-2s} r^{d - 1} \,dr \\
& \leq c N^{d - 2s} = c n^{1 - 2s/d}
\end{align*}
where the last inequality holds since $2s > d$, and $c$ is a constant which depends only on $d$. By plugging back in to~\eqref{eqn:grid_sobolev_approximation_error_pf1} we obtain
\begin{equation*}
\max_{k \in [K]^d}\abs{\theta_k - \wt{\theta}_k} \leq   c n^{1/2 - s/d}
\end{equation*}
and the claim then follows from~\eqref{eqn:grid_sobolev_approximation_error_pf2}.

\subsection{Proof of Lemma~\ref{lem:grid_sobolev_approximation_error_2}}
As shown in the proof of Lemma~\ref{lem:grid_sobolev_approximation_error}, 
\begin{equation}
\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,G}(f)}_n^2 \leq \sqrt{2} \norm{f}_{\Leb^2([0,1]^d)} \left(\sum_{k \in [K]^d} (\theta_k - \wt{\theta}_k)^2\right)^{1/2} + \sum_{k \in \mathbb{N}^d - [K]^d} \theta_k^2 \label{eqn:grid_sobolev_approximation_error_2_pf2}
\end{equation}
We rewrite the discrete Fourier coefficients $\wt{\theta}_k$ as a weighted sum of continuum Fourier coefficients $\theta_k$,
\begin{align}
\wt{\theta}_k - \theta_k & = \frac{1}{n} \sum_{i \in [N]^d} \biggl\{\sum_{l \in \Nbb^d} \theta_l \phi_{l}(\wb{x}_i)\biggr\} \phi_k(\wb{x}_i)  - \theta_k \nonumber \\
& = \sum_{l \in \Nbb^d}  \theta_l \biggl\{\frac{1}{n} \sum_{i \in [N]^d} \phi_{l}(\wb{x}_i) \phi_k(\wb{x}_i)\biggr\}  - \theta_k \label{eqn:grid_sobolev_approximation_error_2_pf1}
\end{align}
Using the tensor decomposition of $\phi_k$, we obtain
\begin{align*}
\frac{1}{n} \sum_{i \in [N]^d} \phi_{l}(\wb{x}_i) \phi_k(\wb{x}_i) & = \frac{1}{n} \sum_{i \in [N]^d} \biggl\{\prod_{j = 1}^{d} \phi_{l_j}(\wb{x}_{ij}) \phi_{k_j}(\wb{x}_{ij}) \biggr\} \\
& = \frac{1}{n} \prod_{j = 1}^{d} \biggl\{\sum_{i_j = 1}^{N}  \phi_{l_j}(\wb{x}_{ij}) \phi_{k_j}(\wb{x}_{ij})  \biggr\} \\
& = \prod_{j = 1}^{d} \Bigl(\delta_{l_j k_j} + \Delta_{l_j k_j}^{(N)}\Bigr)
\end{align*}
where the last line follows from Lemma~\ref{lem:polyak90} (a restatement of results of \textcolor{red}{(Polyak 90)}) and $\Delta_{l_jk_j}$ is as in that Lemma. Inserting this into~\eqref{eqn:grid_sobolev_approximation_error_pf2} gives
\begin{align*}
\abs{\wt{\theta}_k - \theta_k} & = \abs{\sum_{l \in \Nbb^d}  \theta_l \prod_{j = 1}^{d}\Bigl( \delta_{l_j k_j} + \Delta_{l_jk_j}^{(N)}\Bigr)  - \theta_k} \\
& \overset{(i)}{=} \abs{\sum_{l \in \Nbb^d \setminus [N]^d}  \theta_l \prod_{j = 1}^{d}\Bigl( \delta_{l_j k_j} + \Delta_{l_jk_j}^{(N)}\Bigr)} \\
& \overset{(ii)}{\leq} \sum_{l \in \Nbb^d \setminus [N]^d} \prod_{j = 1}^{d} \varepsilon_{l_j}\Bigl(\delta_{l_j k_j} + \abs{\Delta_{l_jk_j}^{(N)}}\Bigr) \\
& \overset{(iii)}{=} \sum_{b \in \{0,1\}^d: \abs{b} \geq 1}\prod_{j: b_j = 0} \varepsilon_{k_j}\prod_{j: b_j = 1} \biggl(\sum_{m = N + 1}^{\infty} \varepsilon_{m} \abs{\Delta_{mk_j}^{(N)}}\biggr) \\
& \overset{(iv)}{\leq} c \sum_{b \in \{0,1\}^d : \abs{b} \geq 1} N^{-\abs{b}} \prod_{j: b_j = 0} \varepsilon_{k_j} 
\end{align*}
where $(i)$ and $(iv)$ follow from Lemma~\ref{lem:polyak90}, $(ii)$ follows from the condition~\eqref{asmp:fourier_coefficient_decay}, and $(iii)$ follows since for each index $l_j$ at most one of $\delta_{l_j k_j}$ or $\Delta_{l_j k_j}^{(N)}$ can be non-zero, and $\delta_{l_j k_j}$ must equal $0$ for at least one $j = 1,\ldots,d$.

We can now upper bound the sum of squared differences between discrete Fourier coefficients and Fourier coefficients,
\begin{align*}
\sum_{k \in [K]^d} (\wt{\theta}_k - \theta_k)^2 & \leq c \sum_{k \in [K]^d} \biggl(\sum_{b \in \{0,1\}^d : \abs{b} \geq 1} N^{-\abs{b}} \prod_{j: b_j = 0} \varepsilon_{k_j} \biggr) \\
& \leq c \sum_{k \in [K]^d} \sum_{b \in \{0,1\}^d : \abs{b} \geq 1} N^{-2\abs{b}} \prod_{j: b_j = 0} \varepsilon_{k_j}^2 \\
& \leq c \sum_{b \in \{0,1\}^d : \abs{b} \geq 1} N^{-2\abs{b}} \sum_{k \in [K]^d} \prod_{j: b_j = 0} \varepsilon_{k_j}^2 \\
& \overset{(v)}{\leq} c \sum_{b \in \{0,1\}^d : \abs{b} \geq 1} N^{-2\abs{b}} \norm{f}_{\Leb^2}^{2(d - \abs{b})/d} K^{\abs{b}} \\
& \overset{(vi)}{\leq} c \norm{f}_{\Leb^2}^2 \Bigl( \frac{K}{N^2 \norm{f}_{\Leb^2}^{2/d}} + \frac{K^d}{N^{2d} \norm{f}_{\Leb^2}^{2}} \Bigr)
\end{align*}
where $(v)$ follows from condition~\eqref{asmp:fourier_coefficient_decay}, $(vi)$ since the summand achieves its maximum either when $b = 1$ or $b = d$, and in the above expression $c \geq 0$ is a constant which changes from line to line. Plugging back into~\eqref{eqn:grid_sobolev_approximation_error_pf2} implies Lemma~\ref{lem:grid_sobolev_approximation_error_2}.

\subsection{Proof of Lemma~\ref{lem:grid_sobolev_approximation_error_3}}
As shown in the proof of Lemma~\ref{lem:grid_sobolev_approximation_error}, 
\begin{equation}
\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,G}(f)}_n^2 \leq \sqrt{2} \norm{f}_{\Leb^2([0,1]^d)} \left(\sum_{k \in [K]^d} (\theta_k - \wt{\theta}_k)^2\right)^{1/2} + \sum_{k \in \mathbb{N}^d - [K]^d} \theta_k^2 \label{eqn:grid_sobolev_approximation_error_3_pf1}
\end{equation}

By Lemma~\ref{lem:polyak90}, we have
\begin{align*}
\abs{\wt{\theta}_k - \theta_k} & = \abs{\sum_{l \in \Nbb^d}  \theta_l \prod_{j = 1}^{d}\Bigl( \delta_{l_j k_j} + \Delta_{l_jk_j}^{(N)}\Bigr)  - \theta_k} \\
& = \abs{\sum_{l \in \Nbb^d \setminus [N]^d}  \theta_l \prod_{j = 1}^{d}\Bigl( \delta_{l_j k_j} + \Delta_{l_jk_j}^{(N)}\Bigr)} \\
& = \sum_{b \in \{0,1\}^d: \abs{b} < N} \sum_{l \in I_b(k)} \abs{\theta_l} \prod_{j: b_j = 1}^{d} \abs{\Delta_{l_jk_j}^{(N)}}
\end{align*}
and therefore, upper bounding $l_2$ norm by $l_1$ norm and exchanging sums, we have
\begin{align*}
\biggl(\sum_{k \in [K]^d} (\wt{\theta}_k - \theta_k)^2\biggr)^{1/2} & \leq \sum_{k \in [K]^d} \abs{\wt{\theta}_k - \theta_k} \\
& \leq \sum_{b \in \{0,1\}^d: \abs{b} < N} \sum_{k \in [K]^d}  \sum_{l \in I_b(k)} \abs{\theta_l} \prod_{j: b_j = 1}^{d} \abs{\Delta_{l_jk_j}^{(N)}}.
\end{align*}
In Lemma~\ref{lem:double_counting}, we establish that the following upper bound holds for each $b \in \{0,1\}^d$ such that $\abs{b} < N$:
\begin{equation*}
\sum_{k \in [K]^d}  \sum_{l \in I_b(k)} \abs{\theta_l} \prod_{j: b_j = 1}^{d} \abs{\Delta_{l_jk_j}^{(N)}} \leq 2^d \sum_{l \in \Nbb^d - [N]^d} \abs{\theta_l}
\end{equation*}
and therefore,
\begin{align*}
\biggl(\sum_{k \in [K]^d} (\wt{\theta}_k - \theta_k)^2\biggr)^{1/2} & \leq 4^d \sum_{l \in \mathbb{N}^d - [N]^d} \abs{\theta_l} \\
& \leq 4^d \biggl(\sum_{l \in \mathbb{N}^d - [N]^d} \abs{\theta_l}^2 \abs{l}^{2s}\biggr)^{1/2} \biggl(\sum_{l \in \mathbb{N}^d - [N]^d} \abs{l}^{-2s}\biggr)^{1/2}  \\
& \leq 4^d L^{1/2} \biggl(\sum_{l \in \mathbb{N}^d - [N]^d} \abs{l}^{-2s}\biggr)^{1/2}.
\end{align*}
Note that this upper bound~\eqref{eqn:grid_sobolev_approximation_error_pf1}, which upper bounds the maximum over $k \in [K]^d$ rather than the sum over all $k \in [K]^d$ by the same term, and therefore incurs an additional loss of $K^{d/2}$. Continuing as in the proof of Lemma~\ref{eqn:grid_sobolev_approximation_error} yields the claimed result.

\subsection{Proof of Lemma~\ref{lem:grid_sobolev_approximation_error_4}}
This proof follows the along the same lines as the proof of Lemma~\ref{lem:grid_sobolev_approximation_error_3}, replacing functions $\phi$ in the Fourier basis by functions $\varphi$ in the modified Fourier basis when necessary.

\section{Supporting Theory}
\begin{lemma}[Restatement and Extension of Lemma 1 of \textcolor{red}{(Polyak 90)}]
	\label{lem:polyak90}
	\begin{equation*}
	\frac{1}{n}\sum_{i = 1}^{n} \phi_{k}(\wb{x}_i)\phi_{l}(\wb{x}_i) = \delta_{kl} + \Delta_{kl}^{(n)}
	\end{equation*}
	where $\Delta_{kl}^{(n)} = 0$ if $k,l \leq n$ or $k$ and $l$ are of different parities, and otherwise
	\begin{equation*}
	\abs{\Delta_{kl}} \leq \sum_{m = 1}^{\infty} \delta_{\abs{[k/2] - [l/2]} = mn} + \delta_{\abs{[k/2] + [l/2]} = mn}
	\end{equation*}

	The same statement holds true when replacing $\phi_k$ by $\varphi_k$ and $\phi_l$ by $\varphi_l$, and using symmetric grid points $\wb{X}_i$.
\end{lemma}
\begin{proof}
	We prove only the statement involving $\varphi_k$ and $\varphi_l$. Note that for $k = 1,\ldots,n$, and $i = 1,\ldots,n$,
	\begin{equation*}
	v_{k,i} = \frac{1}{\sqrt{n}}\varphi_k(\wb{x}_i)
	\end{equation*}
	and $\Delta_{kl}^{(n)} = 0$ for $1 \leq k,l \leq n$ immediately follows.

	By restricting the basis $\varphi_k$ to the nodes $\wb{x}_i$ we induce periodicity. In particular, for $k = mn + (j + 1)$ where $m$ and $j$ are non-negative integers and $0 \leq j < n$, we have
	\begin{equation*}
	\varphi_k(\wb{x}_i) \propto
	\begin{cases*}
	\varphi_{j + 1}(\wb{x}_i),& ~~\textrm{if $m \mod 2 = 0$} \\
	\varphi_{n - j}(\wb{x}_i),& ~~\textrm{if $m \mod 2 = 1$}
	\end{cases*}
	\end{equation*}
	Writing $l = m'n + j' + 1$, we therefore see that $\dotp{\varphi_k}{\varphi_l}_n \in \{-1,+1\}$ if and only if (a) $m \mod 2 = m' \mod 2$ and $j = j'$, or (b) $m \mod 2 \neq m' \mod 2$ and $j + 1 = n - j'$; otherwise, $\dotp{\varphi_k}{\varphi_l}_n = 0$. Condition (a) is equivalent to $|[k/2] - [l/2]| = mn$ for some $m \in \mathbb{N}$, and Condition (b) is equivalent to $[k/2] + [m/2] = mn$ for some $m \in \Nbb$, thus proving the claim.

\end{proof}

\begin{lemma}
If $(\theta_k) \in \ell^2(\Nbb)$ satisfies~\eqref{asmp:fourier_coefficient_decay}, then additionally
\begin{equation*}
\max_{1 \leq k \leq n} \biggl(\sum_{l = n + 1}^{\infty} \theta_l \Delta_{kl}\biggr)^2 \leq \frac{C'}{n^2}
\end{equation*}
for some $C'$ which depends only on $C$.
\end{lemma}

For $b \in \{0,1\}^d$, let $I_b(k) = \set{l \in \Nbb^d: l_ib_i = k_ib_i~~\textrm{for each}~~i = 1,\ldots,d}$.

\begin{lemma}
	\label{lem:double_counting}
	For any $\theta: \mathbb{N}^d \to \Reals^+$, any $K \leq N$, and any $b \in \{0,1\}^d$ such that $\abs{b} < N$, 
	\begin{equation}
	\label{eqn:double_counting}
	\abs{\sum_{k \in [K]^d} \sum_{l \in I_b(k)} \theta_l \prod_{j:b_j = 0} \Delta_{l_jk_j}^{(N)}} \leq 2^d \sum_{\mathbb{N}^d -[N]^d} \abs{\theta_l}
	\end{equation}
\end{lemma}
\begin{proof}
	Without loss of generality, suppose $b_1 = \ldots = b_{\abs{b}} = 1$, and $b_{\abs{b} + 1} = \ldots = b_{d} = 0$. We can upper bound the left hand side of ~\eqref{eqn:double_counting} as 
	\begin{equation*}
	\abs{\sum_{k \in [K]^d} \sum_{l \in I_b(k)} \theta_l \prod_{j:b_j = 0} \Delta_{l_jk_j}^{(N)}} \leq \sum_{k_1,\ldots,k_{\abs{b}} = 1}^{K} \sum_{k_{\abs{b} + 1},\ldots,k_{d} = 1}^{K} \sum_{l_{\abs{b} + 1}, \ldots, l_d = N + 1}^{\infty} \abs{\theta_{k_1\ldots k_{\abs{b}} l_{\abs{b} + 1} \ldots l_d}} \prod_{j = \abs{b} + 1}^{d} \abs{\Delta_{l_jk_j}^{(N)}}
	\end{equation*}
	and it is therefore sufficient to show that
	\begin{equation}
	\label{eqn:double_counting_pf1}
	\sum_{k_{\abs{b} + 1},\ldots,k_{d} = 1}^{K} \sum_{l_{\abs{b} + 1}, \ldots, l_d = N + 1}^{\infty} \abs{\theta_{k_1\ldots k_{\abs{b}} l_{\abs{b} + 1} \ldots l_d}} \prod_{j = \abs{b} + 1}^{d} \abs{\Delta_{l_jk_j}^{(N)}} \leq \sum_{l_{\abs{b} + 1},\ldots,l_{d} = N + 1}^{\infty} \abs{\theta_{k_1\ldots k_{\abs{b}} l_{\abs{b} + 1} \ldots l_d}}
	\end{equation}
	Using the upper bound on $\Delta_{kl}$ given by Lemma~\ref{lem:double_counting}, for a given $k_{\abs{b} + 1},\ldots,k_d$ we have
	\begin{align}
	\sum_{l_{\abs{b} + 1}, \ldots, l_d = N + 1}^{\infty} & \abs{\theta_{k_1\ldots k_{\abs{b}} l_{\abs{b} + 1} \ldots l_d}} \prod_{j = \abs{b} + 1}^{d} \abs{\Delta_{l_jk_j}^{(N)}} \nonumber \\
	& \leq \sum_{l_{\abs{b} + 1}, \ldots, l_d = N + 1}^{\infty} \abs{\theta_{k_1\ldots k_{\abs{b}} l_{\abs{b} + 1} \ldots l_d}} \sum_{m_{\abs{b} + 1,\ldots, m_d} = 1}^{\infty} \prod_{j = \abs{b} + 1}^{d} \delta_{\abs{[k_j/2] - [l_j/2]} = m_jN} + \delta_{\abs{[k_j/2] + [l_j/2]} = m_jN} \nonumber \\
	& = \sum_{m_{\abs{b} + 1,\ldots, m_d} = 1}^{\infty} \sum_{l_{\abs{b} + 1}, \ldots, l_d = N + 1}^{\infty} \abs{\theta_{k_1\ldots k_{\abs{b}} l_{\abs{b} + 1} \ldots l_d}} \prod_{j = \abs{b} + 1}^{d} \delta_{\abs{[k_j/2] - [l_j/2]} = m_jN} + \delta_{\abs{[k_j/2] + [l_j/2]} = m_jN} \nonumber \\
	& = \sum_{m_{\abs{b} + 1,\ldots, m_d} = 1}^{\infty} \sum_{l \in L(m,k)} \abs{\theta_l}. \label{eqn:double_counting_pf2}
	\end{align}
	In the above lines, the sums are always over $l_j$ such that $l_j$ and $k_j$ have the same parity, since otherwise $\Delta_{l_j k_j}^{(N)} = 0$ by Lemma~\ref{lem:polyak90}, and $L(m,k) := \cap_{j = 1}^{d} S_j(m,k) \cap \cap_{j = 1}^{d} R_j(l)$ for
	\begin{align*}
	S_j(m,k) & = 
	\begin{cases*}
	\Bigl\{l_j \in \mathbb{N}: l_j = k_j\Bigr\}, & ~\textrm{for $j = 1,\ldots,\abs{b}$} \\
	\Bigl\{l_j \in \mathbb{N} \setminus [N]: [l_j/2] - [k_j/2] = m_j N \Bigr\} \cup \Bigl\{l_j \in \mathbb{N} \setminus{N}: [l_j/2] + [k_j/2] = m_j N \Bigr\}, & ~\textrm{for $j = \abs{b} + 1,\ldots,d$}.
	\end{cases*} \\
	R_j(k) & = \Bigl\{l_j \in \mathbb{N}: l_j \mod 2 = k_j \mod 2\Bigr\}
	\end{align*}
	The key point is that for any distinct $k_j,k_j' \in [K]$ the sets $S_j(m,k)$ and $S_j(m,k')$ are disjoint unless (a) $[k_j/2] = [k_j'/2]$, in which case $R_j(k)$ and $R_j(k')$ are disjoint since $k_j$ and $k_j'$ are of different parity, or (b) $[k_j/2] = N - [k_j/2]$ and $k_j,k_j'$ are of the same parity. Therefore, for any $k \in [K]^d$ and $m \in \mathbb{N}^d$, there are at most $2^{d - \abs{b}} < 2^d$ choices of $k'$ such that $L(m,k)$ and $L(m,k')$ are not disjoint. Moreover, for any $m \in \mathbb{N}^d$ the inclusion $l_j \in S_j(m,k)$ implies that $\abs{l_j} \geq N + 1$. As a result,
	\begin{equation*}
	\bigcup_{k_{\abs{b} + 1},\ldots,k_d = 1}^{N}\Bigl\{\bigcup_{m \in \mathbb{N}^d} L(m,k)\Bigr\} \subseteq \bigcup_{l_{\abs{b} + 1},\ldots, l_{d} = N}^{\infty} (k_1,\ldots,k_{\abs{b}},l_{\abs{b} + 1},\ldots,l_d)
	\end{equation*} Plugging~\eqref{eqn:double_counting_pf2} back into the left hand side of~ \eqref{eqn:double_counting_pf1}, we have
	\begin{align*}
	\sum_{k_{\abs{b} + 1},\ldots,k_{d} = 1}^{K} \sum_{l_{\abs{b} + 1}, \ldots, l_d = N + 1}^{\infty} \abs{\theta_{k_1\ldots k_{\abs{b}} l_{\abs{b} + 1} \ldots l_d}} \prod_{j = \abs{b} + 1}^{d} \abs{\Delta_{l_jk_j}^{(N)}} & \leq \sum_{k_{\abs{b} + 1},\ldots,k_{d} = 1}^{K} \sum_{m_{\abs{b} + 1,\ldots, m_d} = 1}^{\infty} \sum_{l \in L(m,k)} \abs{\theta_l} \\
	& \leq 2^d \sum_{l_{\abs{b} + 1},\ldots,l_{d} = N + 1}^{\infty} \abs{\theta_{k_1\ldots k_{\abs{b}} l_{\abs{b} + 1} \ldots l_d}}
	\end{align*}
	and the proof~\eqref{eqn:double_counting_pf1}--and therefore the proof the Lemma--is complete.
\end{proof}

\subsection{Equivalence between Sobolev classes}
\label{subsec:sobolev_class_equivalence}

\textcolor{red}{Work through details.}

\end{document}

### 2020-03-08 notes/weekly_notes/3.5.20.NotesForWeek.tex
updated weekly notes with tighter grid approximation error bound

@@
 \phi_1(x) & = 1 \\
-\phi_{2k}(x) & = \cos(2k\pi x),~~ k \in \mathbb{N} \\
-\phi_{2k + 1}(x) & = \sin(2k\pi x),~~ k \in \mathbb{N}
+\phi_{2k}(x) & = \sqrt{2}\cos(2k\pi x),~~ k \in \mathbb{N} \\
+\phi_{2k + 1}(x) & = \sqrt{2}\sin(2k\pi x),~~ k \in \mathbb{N}
 \end{align*}
@@
 
-It is possible to improve on Lemma~\ref{lem:grid_sobolev_approximation_error} under \textcolor{red}{(relatively mild)} additional assumptions.
+It is possible to improve on Lemma~\ref{lem:grid_sobolev_approximation_error} under additional assumptions.
 Specifically, we assume the existence of $\varepsilon_j$ satisfying $j \varepsilon_j$ is monotone non-increasing in $j$, such that for every $f \in H_{\textrm{per}}^{s}([0,1]^d)$, $f = \sum_{k \in \mathbb{N}^d} \theta_k \phi_k$,
@@
 for some constant $C < \infty$.
-
-\textcolor{red}{(TODO:)} {Unfortunately these additional assumptions are not mild at all, but instead relatively close to assuming the function is $s$-Holder.}
 
@@
 
-Now we will attempt to find a sharper bound on discretization error than Lemma~\ref{lem:grid_sobolev_approximation_error}, but that requires less severe assumptions than Lemma~\ref{lem:grid_sobolev_approximation_error_2}.
+Unfortunately the additional assumption~\eqref{asmp:fourier_coefficient_decay} is not mild at all, but instead relatively close to assuming the function is $s$-Holder.
+Finally, we state a sharper bound on discretization error than Lemma~\ref{lem:grid_sobolev_approximation_error}, but that requires less severe assumptions than Lemma~\ref{lem:grid_sobolev_approximation_error_2}.
 \begin{lemma}
@@
 	\label{lem:grid_sobolev_approximation_error_4}
-	Suppose $2s < d$ and $f = \sum_{k \in \mathbb{N}^k} \theta_k \varphi_k$ satisfies $f \in \wt{H}_{\textrm{per}}^{s}([0,1]^d;L)$. Then
+	Suppose $2s < d$ and $f = \sum_{k \in \mathbb{N}^k} \theta_k \varphi_k$ satisfies $f \in \wt{H}^{s}([0,1]^d;L)$. Then
 	\begin{equation*}
@@
 	\end{equation*}
+\end{lemma}
+
+So we've done a lot of interesting work, but we only obtained a sufficient bound on the approximation error when $n^{1/2 - s/d} = o(\norm{f}_2)$, which is still not good enough for our purposes.
+In the following Lemma, we state a tighter bound on approximation error when $d = 1$.
+\begin{lemma}
+	\label{lem:grid_sobolev_approximation_error_5}
+	Suppose $2s > 1$.
+Then there exists a constant $c$ which depends only on $s$ such that for any $f = \sum_{k = 1}^{\infty} \theta_k \varphi_k \in \wt{H}_{\textrm{per}}^{s}([0,1]^d;L)$, the following bound holds:
+	\begin{equation*}
+	\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,\wb{G}_1}(f)}_{n}^2 \leq c L^{1/2} \norm{f}_{\Leb^2} n^{-s} + \sum_{k = \kappa + 1}^{\infty} \theta_k^2
+	\end{equation*} 
+\end{lemma}
+
+We extend the results of Lemma~\ref{lem:grid_sobolev_approximation_error_5} to hold for all $d \geq 1$ in Lemma~\ref{lem:grid_sobolev_approximation_error_6}.
+Although Lemma~\ref{lem:grid_sobolev_approximation_error_6} strictly dominates Lemma~\ref{lem:grid_sobolev_approximation_error_5}, we keep the latter around since its proof is somewhat easier to follow.
+\begin{lemma}
+	\label{lem:grid_sobolev_approximation_error_6}
+	Suppose $f = \sum_{k \in \Nbb^d} \theta_k \varphi_k$ satisfies $f \in \wt{H}^s([0,1]^d;L)$ for some $2s > d$. Then, there exists a constant $c$ which depends only on $s$ and $d$ such that
+	\begin{equation}
+	\label{eqn:grid_sobolev_approximation_error_6}
+	\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,\wb{G}_d}(f)}_{n}^2 \leq c L^{1/2} \norm{f}_{\Leb^2} n^{-s/d} + \sum_{k \in \Nbb \setminus [K]^d} \theta_k^2
+	\end{equation}
 \end{lemma}
@@
 
+\subsection{Proof of Lemma~\ref{lem:grid_sobolev_approximation_error_5}}
+Recall that the eigenvectors of the grid graph $v_k := v_k(\wb{G})$ satisfy
+\begin{equation*}
+v_{k}(\wb{x}_i) = \frac{1}{\sqrt{n}} \varphi_{k}(\wb{x}_i)
+\end{equation*}
+for each $i \in [n]$ and $k \in \mathbb{N}$. Therefore we may rewrite $\Pi_{\kappa,\wb{G}}$ as a function of the cosine Fourier basis,
+\begin{align*}
+\Pi_{\kappa,\wb{G}}(f) & = \frac{1}{n} \sum_{k = 1}^{\kappa} \Bigl\{\sum_{i \in [n]} f(\wb{x}_i) \varphi_k(\wb{x}_i)\Bigr\} \varphi_k \\
+& :=  \sum_{k = 1}^{\kappa} \wt{\theta}_k \varphi_k
+\end{align*}
+where we interpret the right hand side as a function over $\wb{X}$. Since $(\varphi_k)$ is an $L_2(\wb{X})$ orthonormal sequence, we obtain
+\begin{align*}
+\norm{\Pi_{\kappa,\wb{G}}(f)}_n^2 & = \sum_{k = 1}^{\kappa} \wt{\theta}_k^2 \\ 
+& = \sum_{k = 1}^{\kappa} {\theta}_k^2 + \sum_{k = 1}^{\kappa} \wt{\theta}_k^2 - \theta_k^2 \\
+& = \norm{f}_{\Leb^2([0,1]^d)}^2 - \sum_{k = \kappa + 1}^{\infty} \theta_k^2 + \sum_{k = 1}^{\kappa} \wt{\theta}_k^2 - \theta_k^2
+\end{align*}
+or equivalently
+\begin{equation*}
+\norm{f}_{\Leb^2([0,1]^d)}^2 - \norm{\Pi_{\kappa,\wb{G}}(f)}_n^2 = \sum_{k = 1}^{\kappa} \theta_k^2 - \wt{\theta}_k^2 + \sum_{k = \kappa + 1}^{\infty} \theta_k^2
+\end{equation*}
+The second term in the above equation is exactly the second term in~\eqref{eqn:grid_sobolev_approximation_error}, and so we focus our attention on the first term, which represents discretization error.
+Applying the parallelogram law and the Cauchy-Schwarz inequality gives
+\begin{equation}
+\label{eqn:grid_sobolev_approximation_error5_pf2}
+\begin{aligned}
+\sum_{k = 1}^{\kappa} \theta_k^2 - \wt{\theta}_k^2 & \leq \sum_{k = 1}^{\kappa} \max\Bigl\{(\theta_k + \wt{\theta}_k)(\theta_k - \wt{\theta}_k),0\Bigr\} \\
+& \leq \left(\sum_{k = 1}^{\kappa} \max\Bigl\{(\theta_k + \wt{\theta}_k)^2,0\Bigr\}\right)^{1/2} \left(\sum_{k = 1}^{\kappa} (\theta_k - \wt{\theta}_k)^2\right)^{1/2} \\
+& \leq \sqrt{2} \norm{f}_{\Leb^2([0,1])} \left(\sum_{k = 1}^{\kappa} (\theta_k - \wt{\theta}_k)^2\right)^{1/2}
+\end{aligned}
+\end{equation}
+
+It is known that the error $\theta_k - \wt{\theta}_k$ is due to the aliasing phenomenon, where high frequency sinusoidal waves agree with low frequency waves at evenly spaced points.
+This phenomenon has a very precise periodic nature -- expressed in Lemma~\ref{lem:alias}--and we use this Lemma to get the upper bound
+\begin{align*}
+\abs{{\theta}_k - \wt{\theta}_k} & = \abs{\sum_{\ell = 1}^{\infty} \theta_{\ell} \frac{1}{n} \sum_{i = 1}^{n} \varphi_{\ell}(\wb{x}_i) \varphi_{k}(\wb{x}_i)} \\
+& = \sum_{m = 1}^{\infty} \abs{\theta_{2mn + k}} + \sum_{m = 0}^{\infty}\abs{\theta_{2(m + 1)n - (k - 1)}}.
+\end{align*}
+Since $\theta$ belongs to the Sobolev ellipsoid $\Theta(s,L)$, by applying Cauchy-Schwarz we can convert the above two tail sums to bounds which explicitly depend on $n$.
+Applying this logic to the first tail sum gives us
+\begin{align}
+\sum_{m = 1}^{\infty} \abs{\theta_{2mn + k}} & \leq \biggl(\sum_{m = 1}^{\infty} \bigl(a_{2mn + k}\theta_{2mn + k}\bigr)^2\biggr)^{1/2} \biggl(\sum_{m = 1}^{\infty} \bigl(a_{2mn + k})^{-2}\biggr)^{1/2}\nonumber \\
+& \leq \biggl(\sum_{m = 1}^{\infty} \bigl(a_{2mn + k}\theta_{2mn + k}\bigr)^2\biggr)^{1/2} \biggl( \sum_{m = 1}^{\infty} (2mn)^{-2s}\biggr)^{1/2} \nonumber \\
+& \leq \biggl(\sum_{m = 1}^{\infty} \bigl(a_{2mn + k}\theta_{2mn + k}\bigr)^2\biggr)^{1/2} \biggl( \Bigl(1 + \int_{1}^{\infty} x^{-2s} \,dx \Bigr) \biggr)^{1/2} (2n)^{-s}   \nonumber \\
+& = \biggl(\sum_{m = 1}^{\infty} \bigl(a_{2mn + k}\theta_{2mn + k}\bigr)^2\biggr)^{1/2} \biggl( \Bigl(1 + \frac{1}{2s - 1} \Bigr) \biggr)^{1/2} (2n)^{-s}\label{eqn:grid_sobolev_approximation_error5_pf3}
+\end{align}
+where we have used the fact that the Riemann sum of a monotone non-increasing function evaluated at right end points is always no greater than the corresponding integral, and the last equality holds since $2s > 1$.
+By using essentially equivalent reasoning, we obtain a comparable bound on the second tail sum
+\begin{equation*}
+\sum_{m = 0}^{\infty}\abs{\theta_{2(m + 1)n - (k - 1)}} \leq \biggl(\sum_{m = 1}^{\infty} \bigl(a_{2(m + 1)n - (k - 1)}\theta_{2(m + 1)n - (k - 1)}\bigr)^2\biggr)^{1/2} \biggl( \Bigl(2 + \frac{1}{2s - 1} \Bigr) \biggr)^{1/2} (2n)^{-s}
+\end{equation*}
+Thus,
+\begin{align}
+\sum_{k = 1}^{\kappa} \bigl(\theta_k - \wt{\theta}_k\bigr)^2 & \leq c(s) n^{-2s}\sum_{k = 1}^{\kappa} \biggl(\sum_{m = 1}^{\infty} \bigl(a_{2mn + k}\theta_{2mn + k}\bigr)^2 + \sum_{m = 0}^{\infty}\bigl(a_{2(m + 1)n - (k - 1)}\theta_{2(m + 1)n - k + 1} \bigr)^2\biggr) \nonumber \\
+& \leq c(s) n^{-2s} \sum_{j = n + 1}^{\infty} \bigl(\theta_j^2 a_j^2)
+\nonumber \\
+& \leq L c(s) n^{-2s}. \label{eqn:grid_sobolev_approximation_error5_pf4}
+\end{align}
+where the second inequality in the preceding display follows since for any $(m,k) \neq (m',k') \in \mathbb{N} \times [n]$,
+\begin{equation*}
+2mn + k \neq 2m'n + k',~~ 2(m + 1)n - k + 1 \neq 2(m' + 1)n - k' + 1, ~~\textrm{and}~~  2mn + k \neq 2(m' + 1)n - k' + 1
+\end{equation*}
+We plug~\eqref{eqn:grid_sobolev_approximation_error5_pf4} back in to~\eqref{eqn:grid_sobolev_approximation_error5_pf2} to obtain the claim of the Lemma.
+
+\subsection{Proof of Lemma~\ref{lem:grid_sobolev_approximation_error_6}}
+Recall that the eigenvectors of the grid graph $v_k := v_k(\wb{G}_d)$ satisfy
+\begin{equation*}
+v_{k}(\wb{x}_i) = \frac{1}{\sqrt{n}} \varphi_{k}(\wb{x}_i)
+\end{equation*}
+for each $i \in [N]^d$ and $k \in \mathbb{N}^d$. Therefore we may rewrite $\Pi_{\kappa,\wb{G}_d}$ as a function of the tensor product cosine Fourier basis,
+\begin{align*}
+\Pi_{\kappa,\wb{G}_d}(f) & = \frac{1}{n} \sum_{k \in [K]^d} \Bigl\{\sum_{i \in [N]^d} f(\wb{x}_i) \varphi_k(\wb{x}_i)\Bigr\} \varphi_k \\
+& :=  \sum_{k \in [K]^d} \wt{\theta}_k \varphi_k
+\end{align*}
+where we interpret the right hand side as a function over $\wb{X}$. Since $(\varphi_k)$ is an $L_2(\wb{X})$ orthonormal sequence, we obtain
+\begin{align*}
+\norm{\Pi_{\kappa,\wb{G}_d}(f)}_n^2 & = \sum_{k \in [K]^d} \wt{\theta}_k^2 \\ 
+& = \sum_{k \in [K]^d} {\theta}_k^2 + \sum_{k \in [K]^d} \wt{\theta}_k^2 - \theta_k^2 \\
+& = \norm{f}_{\Leb^2([0,1]^d)}^2 - \sum_{k \in \Nbb^d \setminus [K]^d} \theta_k^2 + \sum_{k \in [K]^d} \wt{\theta}_k^2 - \theta_k^2
+\end{align*}
+or equivalently
+\begin{equation*}
+\norm{f}_{\Leb^2([0,1]^d)}^2 - \norm{\Pi_{\kappa,\wb{G}_d}(f)}_n^2 = \sum_{k \in [K]^d} \theta_k^2 - \wt{\theta}_k^2 + \sum_{k \in \Nbb^d \setminus [K]^d} \theta_k^2
+\end{equation*}
+The second term in the above equation is exactly the second term in~\eqref{eqn:grid_sobolev_approximation_error_6}, and so we focus our attention on the first term, which represents discretization error.
+Applying the parallelogram law and the Cauchy-Schwarz inequality gives
+\begin{equation}
+\label{eqn:grid_sobolev_approximation_error6_pf1}
+\begin{aligned}
+\sum_{k = 1}^{\kappa} \theta_k^2 - \wt{\theta}_k^2 & \leq \sum_{k \in [K]^d} \max\Bigl\{(\theta_k + \wt{\theta}_k)(\theta_k - \wt{\theta}_k),0\Bigr\} \\
+& \leq \left(\sum_{k \in [K]^d} \max\Bigl\{(\theta_k + \wt{\theta}_k)^2,0\Bigr\}\right)^{1/2} \left(\sum_{k \in [K]^d} (\theta_k - \wt{\theta}_k)^2\right)^{1/2} \\
+& \leq \sqrt{2} \norm{f}_{\Leb^2([0,1])} \left(\sum_{k \in [K]^d} (\theta_k - \wt{\theta}_k)^2\right)^{1/2}
+\end{aligned}
+\end{equation}
+
+It is known that the error $\theta_k - \wt{\theta}_k$ is due to the aliasing phenomenon, where high frequency sinusoidal waves agree with low frequency waves at evenly spaced points.
+This phenomenon has a very precise periodic nature, as expressed in Lemma~\ref{lem:alias}, which allows us to rewrite $\wt{\theta}_k$.
+Note that for every $\ell \in \mathbb{N}^d$, there exists exactly one $m \in \mathbb{N}_0^d$, $o \in [N]^d$ and $b \in \{0,1\}^d$ such that $\ell = 2mN + (1 - b)o + b(-(o - 1))$ (where for $a,b \in \Reals^d$ we denote $ab := (a_1b_1,\ldots,a_db_d)$).
+Adopting the convention $\theta_{\ell} := 0$ if $\ell$ contains a negative index, we have
+\begin{align*}
+\wt{\theta}_k & = \frac{1}{n} \sum_{i \in [N]^d} f(\wb{x}_i) \varphi_k(\wb{x}_i) \\
+& = \frac{1}{n} \sum_{i \in [N]^d} \Bigl(\sum_{\ell \in \Nbb^d} \theta_{\ell} \varphi_{\ell}(\wb{x}_i) \Bigr) \varphi_k(\wb{x}_i) \\
+& = \sum_{\ell \in \mathbb{N}^d} \theta_{\ell} \Bigl(\frac{1}{n} \sum_{i \in [N]^d} \varphi_{\ell}(\wb{x}_i) \varphi_{k}(\wb{x}_i)\Bigr) \\
+& = \sum_{m \in \mathbb{N}_0^d} \sum_{o \in [N]^d}\sum_{b \in \{0,1\}^d} \theta_{I_N(m,b,o)}  \Bigl(\frac{1}{n} \sum_{i \in [N]^d} \varphi_{I_N(m,b,o)}(\wb{x}_i) \varphi_{k}(\wb{x}_i)\Bigr) \\
+& = \sum_{m \in \mathbb{N}_0^d} \sum_{b \in \{0,1\}^d} \theta_{I_N(m,b,k)} 
+\end{align*}
+where the last equality follows from Lemma~\ref{lem:alias}.
+By applying the Cauchy-Schwarz inequality, and noting that $I_N(m,b,k) \geq mN$ (where inequality between two vectors is interpreted entrywise) we get
+\begin{align*}
+\abs{\wt{\theta}_k - \theta_k} & \leq \sum_{m \in \Nbb_0^d \setminus \{0\}} \sum_{b \in \{0,1\}^d} \abs{\theta_{I_N(m,b,k)}} \\
+& \leq \biggl(\sum_{m \in \Nbb_0^d \setminus \{0\}} \sum_{b \in \{0,1\}^d} \Bigl(\theta_{I_N(m,b,k)}a_{I_N(m,b,k)}\Bigr)^2 \biggr)^{1/2} \biggl(\sum_{m \in \Nbb_0^d \setminus \{0\}} \sum_{b \in \{0,1\}^d} \Bigl(a_{I_N(m,b,k)}\Bigr)^{-2} \biggr)^{1/2} \\
+& \leq 2^{d/2}N^{-s } \biggl(\sum_{m \in \Nbb_0^d \setminus \{0\}} \sum_{b \in \{0,1\}^d} \Bigl(\theta_{I_N(m,b,k)}a_{I_N(m,b,k)}\Bigr)^2 \biggr)^{1/2} \biggl(\sum_{m \in \Nbb_0^d \setminus \{0\}}\sum_{j = 1}^{d} m_j^{-2s} \biggr)^{1/2}
+\end{align*}
+Letting $c(d,s)$ be a finite constant which may change from line to line but can depend only on $d$ and $s$, and letting $\Nbb_2 = \Nbb \setminus \{1\}$, we can upper bound the second sum in the previous display,
+\begin{align*}
+\sum_{m \in \Nbb_0^d \setminus \{0\}}\sum_{j = 1}^{d} m_j^{-2s} & \overset{(i)}{\leq} d^{2(s + 1)} \sum_{m \in \Nbb_0^d \setminus \{0\}} \Bigl(\sum_{j = 1}^{d} m_j^{-2}\Bigr)^s \\
+& \overset{(ii)}{\leq} d^{2(s + 1)} \sum_{\wt{d} = 1}^{d} 2^{\wt{d}} \frac{d!}{\wt{d}!} \sum_{m \in \Nbb_2^{\wt{d}}} \Bigl(\sum_{j = 1}^{d} m_j^{-2}\Bigr)^s \\
+& \overset{(iii)}{\leq} d^{2(s + 1)} \sum_{\wt{d} = 1}^{d} 2^{\wt{d}} \frac{d!}{\wt{d}!} \int_{[1,\infty)^{\wt{d}}} \Bigl(\sum_{j = 1}^{\wt{d}} x_j^{-2}\Bigr)^{s} \,dx_1 \ldots \,dx_{\wt{d}} \\
+& \leq \sum_{\wt{d} = 1}^{d} c(d,s) \int_{\sqrt{\wt{d}}}^{\infty} r^{-2s + d - 1} \,dr \\
+& \overset{(iv)}{\leq} c(d,s) < \infty.
+\end{align*}
+where $(i)$ is an equality when $d = 1$, and otherwise follows from Jensen's inequality when $d \geq 2$ and $s \geq 1$; $(ii)$ follows from specially considering the edge cases where $m_j = 0$ or $m_j = 1$ for $j = \wt{d} + 1,\ldots, d$; $(iii)$ uses the fact that the Riemann sum of a monotone non-increasing function evaluated at right end points is always no greater than the corresponding integral; and $(iv)$ follows since $2s > d$.
+We are now in a position to conclude that
+\begin{align}
+\sum_{k \in [K]^d} (\theta_k - \wt{\theta}_k)^2 & \leq c(d,s) N^{-2s} \sum_{k \in [K]^d} \sum_{m \in \Nbb_0^d \setminus \{0\}} \sum_{b \in \{0,1\}^d} \Bigl(\theta_{I_N(m,b,k)}a_{I_N(m,b,k)}\Bigr)^2 \nonumber \\
+& \leq c(d,s) N^{-2s} \sum_{k \in \Nbb^d \setminus [N]^d} \theta_k^2 a_k^2 \nonumber \\
+& \leq L c(d,s) N^{-2s} \label{eqn:grid_sobolev_approximation_error6_pf2}
+\end{align}
+where the second inequality in the preceding display follows since for any $(m,k,b) \neq (m',k',b')$, $I_N(m,b,k) \neq I_N(m',b',k')$.
+We plug~\eqref{eqn:grid_sobolev_approximation_error6_pf2} back in to~\eqref{eqn:grid_sobolev_approximation_error6_pf1} to obtain the claim of the Lemma.
+
 \section{Supporting Theory}
@@
 \end{lemma}
-\begin{proof}
-	We prove only the statement involving $\varphi_k$ and $\varphi_l$. Note that for $k = 1,\ldots,n$, and $i = 1,\ldots,n$,
+
+The following Lemma is similar to~\eqref{lem:polyak90}, but tweaked to be more useful for our setting.
+It applies to the elements in the cosine basis of period $2$, rather than elements in the Fourier basis of period $1$, and is stated slightly differently.
+\begin{lemma}
+	\label{lem:alias}
+	Let $k \in [n]$ and let $\ell \in \Nbb$. There exists a unique $m \in \Nbb_{0}$, $j \in [n]$ and $b \in \{0,1\}$ such that $\ell = 2mn + (1 - b)j - b(j - 1)$. Then
 	\begin{equation*}
-	v_{k,i} = \frac{1}{\sqrt{n}}\varphi_k(\wb{x}_i)
-	\end{equation*}
-	and $\Delta_{kl}^{(n)} = 0$ for $1 \leq k,l \leq n$ immediately follows.
-
-	By restricting the basis $\varphi_k$ to the nodes $\wb{x}_i$ we induce periodicity. In particular, for $k = mn + (j + 1)$ where $m$ and $j$ are non-negative integers and $0 \leq j < n$, we have
-	\begin{equation*}
-	\varphi_k(\wb{x}_i) \propto
+	\frac{1}{n}\sum_{i = 1}^{n} \varphi_{k}(\wb{x}_i)\varphi_{\ell}(\wb{x}_i) =  
 	\begin{cases*}
-	\varphi_{j + 1}(\wb{x}_i),& ~~\textrm{if $m \mod 2 = 0$} \\
-	\varphi_{n - j}(\wb{x}_i),& ~~\textrm{if $m \mod 2 = 1$}
+	1,& ~~ \textrm{if $j = k$} \\
+	0,& ~~ \textrm{otherwise.}
 	\end{cases*}
 	\end{equation*}
-	Writing $l = m'n + j' + 1$, we therefore see that $\dotp{\varphi_k}{\varphi_l}_n \in \{-1,+1\}$ if and only if (a) $m \mod 2 = m' \mod 2$ and $j = j'$, or (b) $m \mod 2 \neq m' \mod 2$ and $j + 1 = n - j'$; otherwise, $\dotp{\varphi_k}{\varphi_l}_n = 0$.
-Condition (a) is equivalent to $|[k/2] - [l/2]| = mn$ for some $m \in \mathbb{N}$, and Condition (b) is equivalent to $[k/2] + [m/2] = mn$ for some $m \in \Nbb$, thus proving the claim.
+\end{lemma}
 
+\begin{proof}
+	Note that for $k = 1,\ldots,n$, and $i = 1,\ldots,n$,
+	\begin{equation*}
+	v_{k,i} = \frac{1}{\sqrt{n}}\varphi_k(\wb{x}_i).
+	\end{equation*}
+	Additionally, it is easy to check that
+	\begin{equation*}
+	\varphi_{\ell}(\wb{x}_i) \propto
+	\begin{cases*}
+	\varphi_{j}(\wb{x}_i),& ~~\textrm{if $j \in [n]$} \\
+	\varphi_{-(j - 1)}(\wb{x}_i),& ~~\textrm{if $j \in [-(n - 1)]$} \\
+	\varphi_{1}(\wb{x}_i),& ~~\textrm{if $j = 0$.}
+	\end{cases*}
+	\end{equation*}
+	The claim follows since $(v_k)$ is orthonormal in empirical norm.	
 \end{proof}

### 2020-03-12 notes/graph_testing.tex
updated summary documents

@@
 \renewcommand{\bibsection}{\subsubsection*{\bibname}}
+
+\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
+\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
+\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
+\DeclareMathAccent{\wb}{0}{mathx}{"73}
 
@@
 \newcommand{\Rd}{\Reals^d}
+\newcommand{\Nbb}{\mathbb{N}}
 
@@
 We establish some notation that we will use throughout.
-A graph $G = (V,W)$ consists of a collection of vertices $V = \{v_1,\ldots,v_n\}$, and a weight matrix $W \in \Reals^{\abs{V} \times \abs{V}}$ with $ij$th entry $W_{ij}$ encoding affinity between $v_i$ and $v_j$.
+A graph $G = (V,W)$ consists of a collection of vertices $V = \{v_1,\ldots,v_n\}$, and a weight matrix $W \in \Reals^{\abs{V} \times \abs{V}}$ with $(ij)$th entry $W_{ij}$ encoding affinity between $v_i$ and $v_j$.
 We will sometimes write $W_G$ to make it clear which graph a weight matrix is associated with.
@@
 \end{equation*}
-The worst-case risk may be quite close to $1$ unless we enforce some separation between null and alternative spaces. A more realistic measure of performance is therefore
+(Not that the expectation here may taken over the random responses $Y$ when $X$ is assumed to be fixed, or jointly over the randomness of $(X,Y)$ when $X$ is assumed random.) The worst-case risk may be quite close to $1$ unless we enforce some separation between null and alternative spaces.
+A more realistic measure of performance is therefore
 \begin{equation*}
@@
 \end{equation*}
+For a given level $\alpha$ of tolerated error, the minimax critical radius
+\begin{equation*}
+\epsilon^{\star}(\mathcal{H},\alpha) := \inf \Bigl\{\epsilon > 0: \inf_{\phi} R_{\epsilon}(\mathcal{H}) \leq \alpha \Bigr\}
+\end{equation*}
+then provides a lens through which one may examine the hardness of testing over $\mathcal{H}$.
+We will sometimes suppress the dependence of the minimax critical radius on $\alpha$ and simply write $\epsilon^{\star}(\mathcal{H})$.
 
@@
 Letting $\kappa$ be some integer between $1$ and $n$, our test projects $Y$ onto the eigenvectors $v_1(G),\ldots,v_{\kappa}(G)$, and then takes the empirical norm of this projection to be the test statistic.
-Formally,
-\begin{equation*}
-T_{\mathrm{spec}}(G) := \frac{1}{n} \sum_{k = 1}^{\kappa} \Biggl(\sum_{i = 1}^{n} y_i v_{k,i}(G)\Biggr)^2
-\end{equation*}
+Formally, let
+\begin{equation*}
+\Pi_{\kappa,G}(f) := \frac{1}{n}\sum_{k = 1}^{\kappa} \Biggl(\sum_{i = 1}^{n} f(x_i) v_{k,i}(G)\Biggr) v_{k}(G),~~ T_{\mathrm{spec}}(G) := \norm{\Pi_{\kappa,G}(f)}_n^2
+\end{equation*}
+where $\norm{\theta}_n^2 := \frac{1}{n} \sum_{i = 1}^{n} \theta_i^2$ for $\theta \in \Reals^n$.
 Then the test $\phi_{\spec}(G) := \1\{T_{\spec}(G) \geq \tau\}$ rejects the null hypothesis when the test statistic $T_{\mathrm{spec}}(G)$ is greater than a pre-specified cutoff $\tau$.
@@
 Let $R > 0$ and $b \geq 1$ be fixed constants, and $d = 1,2$ or $3$.
-Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p(x)$ is bounded away from zero and infinity,
+Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p(x)$ bounded away from zero and infinity,
 	\begin{equation*}
@@
 We shall discuss testing in the low-smoothness regime shortly.
-First, let's consider the situation when we assume more derivatives worth of smoothness.
+First, let's consider the situation when we are willing assume $f$ possesses additional derivatives bounded in $\Leb^2$ norm.
 
@@
 
-When we additionally believe $H^s(\Xset)$, we know (again from Ingster) that the minimax testing rate is $n^{-4s/(4s + d)}$.
+When we are additionally willing to assume $H^s(\Xset)$, we know (again from \textcolor{red}{Ingster}) that the minimax testing rate is $n^{-4s/(4s + d)}$.
 Now, however, our estimates of the eigenvectors no longer track the derivatives finely enough at the boundaries of $\Xset$ to establish the same minimax rate over all $H^s(\mathcal{X})$.
@@
 	\begin{equation*}
-	n^{-1/(2(s-1) + d)} \leq r(n) \leq n^{-4/((4s + d)(2+d))}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	n^{-1/(2(s-1) + d)} \leq r(n) \leq n^{-4/((4s + d)(2+d))}, ~\kappa = (Rn)^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
 	\end{equation*}
@@
 	\label{eqn:L4_testing_rate}
-	\epsilon^2 > b \cdot n^{-1/2}
+	\epsilon^2 > c^2 b \cdot n^{-1/2}
 	\end{equation}
@@
 	\begin{equation*}
-	\phi_{\mathrm{mean}} = \1\{\frac{1}{n}\sum_{i = 1}^{n} y_i^2 \geq 1\}
+	\phi_{\mathrm{mean}} = \1\Bigl\{\frac{1}{n}\sum_{i = 1}^{n} y_i^2 \geq 1\Bigr\}
 	\end{equation*}
@@
 
+\subsection{Fixed Design}
+Frequently in nonparametric regression problems, it is assumed that the design points are fixed, and, in the simplest case, uniformly spaced on the unit cube.
+For convenience in this setting we let $n = N^d$ for some integer $N$.
+The fixed grid design $\wb{X}$ consists of $n$ total evenly spaced grid points on $[0,1]^d$,
+\begin{equation*}
+\wb{X} := \Bigl\{\Bigl(\frac{2i_1 - 1}{2N},\ldots,\frac{2i_d - 1}{2N}\Bigr): i \in [N]^d\Bigr\}
+\end{equation*}
+Suppose we observe $n$ samples according to regression model
+\begin{equation}
+\label{eqn:grid_regression_model}
+y_i = f(\wb{x}_i) + \varepsilon_i, ~\varepsilon_i \overset{\mathrm{i.i.d}}{\sim} \mathcal{N}(0,1) ~~\textrm{for each $i \in [N]^d$,}
+\end{equation}
+Clearly model~\eqref{eqn:grid_regression_model} bears some similarity with model~\eqref{eqn:regression_random_design_known_variance}, the main difference being that the design $\wb{X}$ is now fixed instead of random.
+
+We study the behavior of $\phi_{\spec}(\wb{G})$, where $\wb{G}$ is the lattice
+\begin{equation*}
+\wb{G} := \bigl(\wb{X},\wb{E}\bigr),~~ \wb{E} := \set{(\wb{x}_i,\wb{x}_j): i,j \in [N]^d, \norm{i - j}_1 = 1},
+\end{equation*}
+a very natural graph to build given our assumed grid design. Note that for a given $d$, we also have the recursive relation
+\begin{equation*}
+\wb{G} := \wb{G}_1 \otimes \ldots \otimes \wb{G}_1
+\end{equation*}
+where the tensor product is taken $d$ total times, and $\wb{G}_1$ is the $1$d lattice (i.e.
+the path) graph on $N$ vertices.
+For this reason it will be convenient to change notation slightly, and index by $d$-tuples of numbers rather than by numbers.
+In particular, we index the eigenvalues and eigenvectors of $L_{\wb{G}}$ as
+\begin{equation*}
+\lambda_k(\wb{G}) = \prod_{j = 1}^{d} \lambda_{k_j}(\wb{G}_1), v_{k,i}(\wb{G}) = \prod_{j = 1}^{d} v_{k_j,i_j}(\wb{G}_1)~~\textrm{for $k,i \in [N]^d$.}
+\end{equation*}
+Then, for a given $\kappa \in [n]$, letting $K = \kappa^{1/d}$, we see that our graph spectral test statistic may be written equivalently as
+\begin{equation*}
+T_{\spec}(\wb{G}) = \frac{1}{n} \sum_{k \in [K]^d} \Biggl(\sum_{i \in [N]^d} y_i v_{k,i}(G)\Biggr)^2
+\end{equation*}
+
+We see the test $\phi_{\spec}(\wb{G})$ achieves the same rate over the Holder ball $C^{1}(\Xset;L)$ when $d \leq 4$, as the test $\phi_{\spec}(G_{n,r})$ does over the Holder ball $C^1(\Xset;L)$.
+\begin{theorem}
+	\label{thm:holder_testing_rate_grid}
+	Let $L > 0$ be a fixed constant, and let $d \leq 4$.
+Suppose we observe data according to the model~\eqref{eqn:grid_regression_model}.
+Then when the test $\phi_{\spec}(\wb{G})$ is performed with the parameter choices
+	\begin{equation*}
+	\kappa = (Ln)^{2d/(4 + d)},~~ \tau(b) = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	\end{equation*}
+	for some $b \geq 1$, the following statement is true: there exists a constant $c$ such that for 
+	\begin{equation}
+	\label{eqn:holder_testing_rate_grid_1}
+	\epsilon^2 \geq c \cdot \max\{L^{d/(4+d)}, L^{d/(4 + d) + 1}\} \cdot b n^{-4/(4 + d)}
+	\end{equation}
+	the worst-case risk of $\phi_{\spec}(\wb{G})$ on $C^{1}(\Xset;L)$ is upper bounded
+	\begin{equation}
+	\label{eqn:holder_testing_rate_grid_2}
+	\mathcal{R}_{\epsilon}\Bigl(\phi_{\spec}(\wb{G});C^{1}(\Xset;L)\Bigr) \leq \frac{4}{b^2} + \frac{8}{b\sqrt{2\kappa}}.
+	\end{equation}
+\end{theorem}
+
+As was the case with the neighborhood graph in the random design setting, the grid Sobolev norm fails to track the higher order derivatives of $f$ near the boundary of $\Xset$.
+We therefore will again insist that $f$ be compactly supported within $\Xset$.
+Under this condition, we find that the test $\phi_{\spec}(\wb{G})$ achieves the same rate over the Holder ball $C_c^{s}(\Xset;L)$ when $4s \geq d$ as did the test $\phi_{\spec}(G_{n,r})$ over the Holder ball $C_c^s(\Xset;L)$.
+
+\begin{theorem}
+	\label{thm:holder_testing_rate_grid_higher_order}
+	Let $L > 0$ be a fixed constant, $s \geq 2$ be a fixed integer, and let $d \leq 4s$.
+Suppose we observe data according to the model~\eqref{eqn:grid_regression_model}.
+Then when the test $\phi_{\spec}(\wb{G})$ is performed with the parameter choices
+	\begin{equation*}
+	\kappa = (Ln)^{2d/(4s + d)},~~ \tau(b) =  \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	\end{equation*} 
+	for some $b \geq 1$, the following statement is true: there exists a constant $c$ such that for any
+	\begin{equation*}
+	\epsilon^2 \geq c L^2 b n^{-4s/(4s + d)},
+	\end{equation*}
+	the worst-case risk of $\phi_{\spec}(\wb{G})$ is upper bounded
+	\begin{equation*}
+	\mathcal{R}_{\epsilon}\Bigl(\phi_{\spec}(\wb{G});C_c^{s}(\Xset;L)\Bigr) \leq \frac{4}{b^2} + \frac{8}{b\sqrt{2\kappa}}.
+	\end{equation*}
+\end{theorem}
+
+Roughly speaking, we have shown that in the high-smoothness regime $4s > d$, testing over Holder classes in the grid design and random design cases is an equally difficult problem from the minimax perspective.
+The same is not true for the corresponding Sobolev spaces.
+When $2s > d$, the Sobolev spaces $H^s(\Xset;L)$ embed compactly into (lower-order) Holder classes, implying that every equivalence class $u \in H^s(\Xset;L)$ contains a favored representer which is bounded and Holder continuous, with bounded Holder norm.
+When $2s < d$, the situation is less nice; the spaces $H^s(\Xset;L)$ now contain (equivalence classes of) bump functions of arbitrarily large height and small width.
+By placing such bump functions at design points, it is possible to make the sampling model~\eqref{eqn:grid_regression_model} contain arbitrarily little information about the overall behavior of the function $f$ over $\Xset$.
+It is not hard to construct a function $f \in H^s(\Xset;L)$ for which, say, $\norm{f}_{\Leb^2(\Xset)} = 1$ yet $f(\wb{x}) = 0$ for all $\wb{x} \in \wb{X}$, implying that the critical radius $\epsilon^{\star}(H^s(\Xset;L)) = \Omega(1)$.
+We note that this is not an issue in the random design setting, because the design points are (randomly) chosen only after the function $f$ is fixed.
+
+\textcolor{red}{(TODO): Make this discussion more rigorous. Perhaps appeal to the concept of precise representatives defined in (Evans and Gariepy).}
+
+The above discussion leaves hope that when $2s > d$, consistent testing might still be possible.
+We can in fact go beyond this and show that the graph spectral test $\phi_{\spec}(\wb{G})$ achieves the same rate over the Sobolev space $\wt{H}^s(\Xset;L)$ as the test $\phi_{\spec}(G_{n,r})$ did over $H_0^s(\Xset;L)$.
+
+\begin{theorem}
+	\label{thm:sobolev_testing_rate_grid}
+	Let $L > 0$ be a fixed constant, $s \geq 1$ be a fixed integer, and let $2s > d$.
+Suppose we observe data according to the model~\eqref{eqn:grid_regression_model}.
+Then when the test $\phi_{\spec}(\wb{G})$ is performed with the parameter choices
+	\begin{equation*}
+	\kappa = n^{2d/(4s + d)},~~ \tau(b) =  \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	\end{equation*} 
+	for some $b \geq 1$, the following statement is true: there exists a constant $c$ such that for any
+	\begin{equation*}
+	\epsilon^2 \geq c L^2 b n^{-4s/(4s + d)},
+	\end{equation*}
+	the worst-case risk of $\phi_{\spec}(\wb{G})$ is upper bounded
+	\begin{equation}
+	\label{eqn:sobolev_testing_rate_grid_1}
+	\mathcal{R}_{\epsilon}\Bigl(\phi_{\spec}(\wb{G});\wt{H}^s(\Xset;L)\Bigr) \leq \frac{4}{b^2} + \frac{8}{b\sqrt{2\kappa}}
+	\end{equation}
+	for all $n$ sufficiently large.
+\end{theorem}
+
+\subsubsection{Low-smoothness regime with fixed design}
+In the low-smoothness regime $d > 4s$, the coupling between empirical norm over $\wb{X}$ and the continuum $\Leb^2$ norm is sufficiently weak that the discretization error can become the dominant source of error, as we show in Lemma~\ref{lem:holder_testing_rate_grid_low_smoothness_lb}.
+
+\begin{lemma}
+	\label{lem:holder_testing_rate_grid_low_smoothness_lb}
+	For any $L > 0$ and any integers $s$ and $d$, there exists a function $g \in C^s(\Xset;L)$ such that
+	\begin{equation*}
+	\norm{g}_{\Leb^2}^2 \geq \frac{L^2}{2^d \pi^{2s} (d + 1)^{2s}}n^{-2s/d} 
+	\end{equation*}
+	but $g(\wb{x}) = 0$ for all $\wb{x} \in \wb{X}$. 
+\end{lemma}
+Lemma~\ref{lem:holder_testing_rate_grid_low_smoothness_lb} implies the critical radius over the Holder classes $C^s(\Xset;L)$ must be at least on the order of $\epsilon^{\star}(C^s(\Xset;L)) \gtrsim n^{-s/d}$.
+When $4s \geq d$, this term is neglible relative to $n^{-2s/(4s + d)}$, and we arrive at the ``usual'' rates of minimax testing we see in Theorems~\ref{thm:holder_testing_rate_grid} and~\ref{thm:holder_testing_rate_grid_higher_order}.
+On the other hand when $4s < d$, this becomes the dominant term, and we can no longer achieve the typical rate $\epsilon^2 \asymp n^{-4s/(4s + d)}$.
+
+In this setting, when we enforce the wider radius $\epsilon^2 \asymp n^{-2s/d}$, Proposition~\ref{prop:holder_testing_rate_grid_low_smoothness_ub} shows that a very simply test has small worst-case risk $\mathcal{R}_{\epsilon}$.
+
+\begin{proposition}
+	\label{prop:holder_testing_rate_grid_low_smoothness_ub}
+	Let $L > 0$ be a fixed constant, $s \geq 1$ a fixed integer, and let $d \geq 4s$.
+Suppose we observe data according to the grid design regression model~\eqref{eqn:grid_regression_model}.
+Then when the test $\phi_{\spec}(\wb{G})$ is performed with the parameter choices
+	\begin{equation*}
+	\kappa = n,~~ \tau(b) = 1 + b\sqrt{\frac{2}{n}}
+	\end{equation*}
+	for some $b \geq 1$, the following statement is true: there exists a constant $c$ such that for any 
+	\begin{equation*}
+	\epsilon^2 \geq c b L^2 n^{-2s/d}
+	\end{equation*}
+	the worst case risk of $\phi_{\spec}(\wb{G})$ is upper bounded
+	\begin{equation*}
+	\mathcal{R}_{\epsilon}(\phi_{\spec}(\wb{G}); C^s(\Xset;L)) \leq \frac{4}{b^2} + \frac{8}{b\sqrt{2n}}
+	\end{equation*}
+\end{proposition}
+
+\begin{itemize}
+	\item The test statistic $T_{\spec}(\wb{G})$ is simply the empirical norm of $Y$. 
+	\item Together, Lemma~\ref{lem:holder_testing_rate_grid_low_smoothness_lb} and Proposition~\eqref{prop:holder_testing_rate_grid_low_smoothness_ub} characterize the minimax rate of the grid design regression testing problem when $4s > d$.
+	\item This minimax rate does not match the upper bound established in Proposition~\ref{prop:L4_testing_rate}---the fixed design problem is harder. 
+	\item Note that the usual minimax rate of estimation error over Holder classes is always $n^{-2s/(2s + d)}$, regardless of the relation between $s$ and $d$, and regardless of whether loss is measured in $\Leb^2$-norm or empirical norm.
+The explanation for this is that the rate $n^{-2s/(2s + d)}$ is always larger than $n^{-2s/d}$, and so the discretization error is never the dominant source of error in estimation.
+This reveals an interesting distinction between the testing and estimation problems, as testing is easy enough in a statistical sense that error incurred by the discrete nature of the problem may be the bottleneck.
+\end{itemize}
+
 \subsection{Analysis}
@@
 \begin{equation}
+\label{eqn:fixed_graph_regression_model}
 y_i = \beta_i + \varepsilon_i,~~\varepsilon_i \sim \mathcal{N}(0,1)
@@
 	Let $1 \leq \kappa \leq n$ be an integer.
+Suppose we observe data according to the model~\eqref{eqn:fixed_graph_regression_model}, and perform the test $\phi_{\spec}(G)$ with $\tau(b) = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}$.
 	\begin{enumerate}
-		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}$ is upper bounded
+		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}(G)$ is upper bounded
 		\begin{equation}
 		\label{eqn:graph_spectral_type_I_error}
-		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{1}{b^2}.
+		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{2}{b^2}.
 		\end{equation}
 		\item \textbf{Type II error:} For any $b \geq 1$ and $\beta$ such that
+		\begin{equation}
+		\label{eqn:fixed_graph_testing_critical_radius_1}
+		\norm{\Pi_{\kappa,G}(\beta)}_n^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}}
+		\end{equation}
+		the Type II error of $\phi_{\spec}(G)$ is upper bounded,
+		\begin{equation}
+		\label{eqn:graph_spectral_type_II_error}
+		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{2}{b^2} + \frac{8}{b\sqrt{2\kappa}}.
+		\end{equation}
+		In particular if
 		\begin{equation}
@@
 		\end{equation}
-		the Type II error of $\phi_{\spec}$ is upper bounded,
-		\begin{equation}
-		\label{eqn:graph_spectral_type_II_error}
-		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{3}{b}.
-		\end{equation}
+		then~\eqref{eqn:fixed_graph_testing_critical_radius_1} and thus~\eqref{eqn:graph_spectral_type_II_error} follow.
 	\end{enumerate}
@@
 
-\paragraph{Lebesgue spaces.}
+Here we formally define, discuss some function spaces and associated norms.
+
+\subsubsection{Lebesgue spaces}
 
@@
 
-\paragraph{Holder spaces.}
+\subsubsection{Holder spaces}
 
@@
 
-\paragraph{Sobolev spaces.}
+\subsubsection{Sobolev spaces}
 
@@
 
-We write $W_0^{s,p}(\Xset)$ for the completion of $C_c^{\infty}(\Xset)$, and let $H_0(\Xset) = W_0^{s,2}(\Xset)$.
+We write $W_0^{s,p}(\Xset)$ for the completion of $C_c^{\infty}(\Xset)$, and let $H_0^s(\Xset) = W_0^{s,2}(\Xset)$.
+
+\paragraph{Defining Sobolev spaces using Fourier series.}
+
+Let the cosine basis on $\Leb^2([0,1])$ be defined as 
+\begin{equation*}
+\varphi_k(x) = 
+\begin{cases*}
+1,~~\textrm{for $k = 1$.} \,
+\sqrt{2}\cos\Bigl(\pi (k - 1) x\Bigr),~~\textrm{for $k \in \Nbb_1$}
+\end{cases*}
+\end{equation*}
+for any $x \in [0,1]$. Let the tensor product cosine basis on $\Leb^2([0,1]^d)$ be defined as
+\begin{equation*}
+\varphi_k(x) = \prod_{j = 1}^{d} \varphi_{k_j}(x_j), ~~\textrm{for $k \in \Nbb^d$},
+\end{equation*}
+for any $x \in [0,1]^d$. Observe that $\varphi_k(\wb{x}_i) = \sqrt{n} v_{k,i}(\wb{G}_d)$ for all $k,i \in [N]^d$.
+
+We let the space 
+\begin{equation*}
+\wt{H}^{s}(\Xset;L) = \Bigl\{ \sum_{k \in \mathbb{N}^d} \theta_k \varphi_k: \sum_{k \in \mathbb{N}^d} \theta_k^2 a_k^{2} \leq L \Bigr\}
+\end{equation*}
+where in the one-dimensional case $a_{k} = (k - 1)^s$ and for general $d$, $a_k = \sqrt{\sum_{j = 1}^{d} a_{k_j}^2}$.
+It is not hard to show that, when $s$ is an integer, $\wt{H}^s([0,1]^d)$ consists of those functions in $H^s([0,1]^d)$ which satisfy some Neumann-type boundary conditions.
+For example when $d = 1$,
+\begin{equation*}
+\wt{H}^s([0,1];L) = \Bigl\{f \in H^{s}([0,1];L \pi^{s}): f^{(\ell)}(0) = f^{(\ell)}(1) = 0,~~\textrm{for all odd integers $0 < \ell < s$} \Bigr \} \supseteq H_0^s([0,1];L'),
+\end{equation*}
+a fact which we prove in Section~\ref{subsec:sobolev_class_equivalence}.
 
@@
 
+We sometimes -- and currently not in a consistent manner -- denote the lattice Laplacian $L_{\wb{G}} =: \wb{L}$.
+
+We write $\mathbb{N}_k$ for the natural numbers excluding $0,\ldots,k - 1$. We write $\mathbb{N}_0$ for the natural numbers including $0$, and $\mathbb{N} = \mathbb{N}_1$.
+
 \section{Proofs}
@@
 \Ebb(T_{\spec}) & = \frac{1}{n}\left(\sum_{k = 1}^{\kappa} \dotp{\beta}{v_k}^2 + \Ebb\bigl( \dotp{\varepsilon}{v_k}^2 + 2 \dotp{\varepsilon}{v_k} \dotp{\beta}{v_k}\bigr)\right) \\
-& = \frac{\kappa}{n} + \frac{1}{n}\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2.
-\end{align*}
-When $\beta = 0$, this equals $\kappa/n$. Otherwise, we have the following lower bound:
+& = \frac{\kappa}{n} + \frac{1}{n}\sum_{k = 1}^{\kappa}\dotp{\beta}{v_k}^2 \\
+& = \frac{\kappa}{n} + \norm{\Pi_{\kappa,G}(\beta)}_n^2.
+\end{align*}
+
+\paragraph{Variance of $T_{\mathrm{spec}}$:}
+Note that $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvector $v_k$ as its $k$th column. Therefore,
+\begin{align*}
+\Var(T_{\spec}) & = \frac{1}{n^2} \Var(y^T V_{\kappa} V_{\kappa}^T y) \\
+& = \frac{1}{n^2} \Var((\beta + \varepsilon)^T V_{\kappa} V_{\kappa}^T (\beta + \varepsilon)) \\
+& = \frac{1}{n^2} \Var(2 \beta^T V_{\kappa} V_{\kappa}^T \varepsilon + \varepsilon^T V_{\kappa} V_{\kappa}^T \varepsilon) \\
+& \leq \frac{1}{n^2}(4 \beta^T V_{\kappa} V_{\kappa}^T \beta + 2\kappa) \\
+& = \frac{1}{n^2}(4 n \norm{\Pi_{\kappa,G}(\beta)}_n^2 + 2 \kappa)
+\end{align*}
+where the last inequality follows from standard properties of the Gaussian distribution, and the last equality follows since $V_{\kappa} V_{\kappa}^T$ is idempotent.
+We now move on to showing the desired inequalities \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.
+Let $t(b) := b\sqrt{\frac{\kappa}{n^2}}$.
+
+\paragraph{Proof of~\eqref{eqn:graph_spectral_type_I_error}:} By Chebyshev's inequality,
+\begin{align*}
+\Pbb_{\beta = 0}\bigl(T_{\spec} \geq \frac{\kappa}{n} + t(b)\bigr)
+& \leq \Pbb_{\beta = 0}\bigl(\abs{T_{\spec} - \frac{\kappa}{n}} \geq t(b)\bigr) \\
+& \leq \frac{\Var_{\beta = 0}(T_{\spec})}{t(b)^2} = \frac{2}{b^2}.
+\end{align*}
+
+\paragraph{Proof of~\eqref{eqn:graph_spectral_type_II_error}:} For simplicity, we introduce the notation
+\begin{equation*}
+\Delta = \norm{\Pi_{\kappa,G}(\beta)}_n^2
+\end{equation*}
+Assumption~\eqref{eqn:fixed_graph_testing_critical_radius} implies $\Delta \geq 2 t(b)$, and we have already shown that $\Ebb_{\beta}(T_{\spec}) = \Delta + \kappa/n$.
+Then another application of Chebyshev's inequality gives us
+\begin{align*}
+\Pbb_{\beta}\bigl(T_{\spec} \leq \frac{\kappa}{n} + t(b)\bigr) & \leq \Pbb_{\beta}\bigl(T_{\spec} - \Ebb_{\beta}(T_{\spec}) \leq t(b) - \Delta \bigr) \\
+& \leq \Pbb_{\beta}\bigl(\abs{T_{\spec} - \Ebb_{\beta}(T_{\spec})} \leq \Delta - t(b) \bigr) \tag{since $\Delta \geq t(b)$}	\\
+& \leq \frac{\Var_{\beta}(T_{\spec})}{(\Delta - t(b))^2} \\
+& \leq 4\frac{\Var_{\beta}(T_{\spec})}{\Delta^2} \tag{since $\Delta \geq 2t(b)$} \\
+& \leq 4\frac{2\kappa/n^2 + 4\Delta/n}{\Delta^2}.
+\end{align*}
+We handle each summand in the numerator separately. For the first term, since $\Delta \geq 2 t(b)$, we have
+\begin{equation}
+\label{eqn:spectral_type_II_error_pf1}
+\frac{2\kappa}{n^2\Delta^2} \leq \frac{1}{2b^2}.
+\end{equation}
+
+For the second term we have
+\begin{equation}
+\label{eqn:spectral_type_II_error_pf2}
+\frac{4}{n\Delta} \leq \frac{2}{nt(b)} = \frac{2}{b\sqrt{2\kappa}}, 
+\end{equation}
+and combining~\eqref{eqn:spectral_type_II_error_pf1} and~\eqref{eqn:spectral_type_II_error_pf2} yields~\eqref{eqn:graph_spectral_type_II_error}.
+
+Finally the following lower bound:
 \begin{align*}
@@
 \end{align*}
-and therefore $\Ebb(T_{\spec}) \geq \kappa/n + n^{-1}(\norm{\beta}_2^2 - \beta^T L^s \beta/\lambda_{\kappa}^s)$.
-
-\paragraph{Variance of $T_{\mathrm{spec}}$:}
-We write $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvector $v_k$ as its $k$th column.
-\begin{align*}
-\Var(T_{\spec}) & = \frac{1}{n^2} \Var(y^T V_{\kappa} V_{\kappa}^T y) \\
-& = \frac{1}{n^2} \Var((\beta + \varepsilon)^T V_{\kappa} V_{\kappa}^T (\beta + \varepsilon)) \\
-& = \frac{1}{n^2} \Var(2 \beta^T V_{\kappa} V_{\kappa}^T \varepsilon + \varepsilon^T V_{\kappa} V_{\kappa}^T \varepsilon) \\
-& \leq \frac{1}{n^2}(4 \beta^T V_{\kappa} V_{\kappa}^T \beta + 2\kappa)
-\end{align*}
-where the last inequality follows from standard properties of the Gaussian distribution.
-We now move on to showing the desired inequalities \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.
-
-\paragraph{Proof of~\eqref{eqn:graph_spectral_type_I_error}:} By Chebyshev's inequality,
-\begin{align*}
-\Pbb_{\beta = 0}\bigl(T_{\spec} \geq \frac{\kappa}{n} + t(b)\bigr)
-& \leq \Pbb_{\beta = 0}\bigl(\abs{T_{\spec} - \frac{\kappa}{n}} \geq t(b)\bigr) \\
-& \leq \frac{\Var_{\beta = 0}(T_{\spec})}{t(b)^2} = \frac{1}{b^2}.
-\end{align*}
-
-\paragraph{Proof of~\eqref{eqn:graph_spectral_type_II_error}:} For simplicity, we introduce the notation
-\begin{equation*}
-\Delta = \frac{\norm{\beta}_2^2}{n} - \frac{\beta^T L^s \beta}{n\lambda_{\kappa}^s}.
-\end{equation*}
-Assumption~\eqref{eqn:fixed_graph_testing_critical_radius} implies $\Delta \geq 2 t(b)$. Then another application of Chebyshev's inequality gives us
-\begin{align*}
-\Pbb_{\beta}\bigl(T_{\spec} \leq \frac{\kappa}{n} + t(b)\bigr) & = \Pbb_{\beta}\bigl(T_{\spec} - \Ebb_{\beta}(T_{\spec}) \leq t(b) - \Delta \bigr) \\
-& \leq \Pbb_{\beta}\bigl(\abs{T_{\spec} - \Ebb_{\beta}(T_{\spec})} \leq \Delta - t(b) \bigr) \tag{since $\Delta \geq t(b)$}	\\
-& \leq \frac{\Var_{\beta}(T_{\spec})}{(\Delta - t(b))^2} \\
-& \leq 4\frac{\Var_{\beta}(T_{\spec})}{\Delta^2} \tag{since $\Delta \geq 2t(b)$} \\
-& \leq 4\frac{2\kappa/n^2 + 4\beta^T V_{\kappa} V_{\kappa}^T \beta /n^2}{\Delta^2}.
-\end{align*}
-We handle each summand in the numerator separately. For the first term, since $\Delta \geq 2 t(b)$, we have
-\begin{equation}
-\label{eqn:spectral_type_II_error_pf1}
-\frac{2\kappa}{n^2\Delta^2} \leq \frac{1}{2b^2}.
-\end{equation}
-
-For the second term, noting that $\Delta = \beta^T V_{\kappa} V_{\kappa}^T \beta/n$, we have
-\begin{align}
-\frac{\beta^T V_{\kappa} V_{\kappa}^T \beta/n^2}{\Delta^2} & = \frac{1}{n\Delta} \nonumber \\
-& \leq \frac{1}{2nt(b)} \nonumber \\
-& = \frac{1}{2b\sqrt{2\kappa}}, \label{eqn:spectral_type_II_error_pf2}
-\end{align}
-and combining~\eqref{eqn:spectral_type_II_error_pf1} and~\eqref{eqn:spectral_type_II_error_pf2} yields~\eqref{eqn:graph_spectral_type_II_error}.
+shows that~\eqref{eqn:fixed_graph_testing_critical_radius} implies~\eqref{eqn:fixed_graph_testing_critical_radius_1}.
 
@@
 	\end{equation}
+	Here $L = L_{G_{n,r}}$ is the Laplacian matrix of the neighborhood graph $G_{n,r}$.
 	\item 
@@
 completing the proof of Theorem~\ref{thm:sobolev_testing_rate} upon proper choice of constants $c_1,c_2$ in that Theorem.
+
+\subsection{Proof of Theorem~\ref{thm:holder_testing_rate_grid}}
+\label{subsec:holder_testing_rate_grid_pf}
+
+We want to apply Lemma~\ref{lem:fixed_graph_testing} to the case where $G = \wb{G}$ and $s = 1$.
+In order to do this, we will need to show that when $r$ and $\kappa$ are appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = \wb{G}$ and $\beta = (f(\wb{x}_i): i \in [N]^d)$.
+In particular, we will show that each of the following statements hold:
+
+\begin{enumerate}[label=(E\arabic*)]
+	\item
+	\label{event:discrete_sobolev_norm_grid}
+	\textbf{Lattice Sobolev norm:}
+	For every $f \in C^1(\Xset)$,
+	\begin{equation}
+	\label{eqn:discrete_sobolev_norm_grid}
+	f^T L f \leq 2d \norm{f}_{C^1(\Xset)}^2 \cdot n^{1 - 2/d}
+	\end{equation}
+	Here $L = L_{\wb{G}}$ is the Laplacian matrix of the lattice graph.
+
+	\item
+	\label{event:eigenvalue_tail_bound_grid}
+	\textbf{Lattice eigenvalue tail bound:}
+	Letting $\mathbf{K} = (K,\ldots,K)$, 
+	\begin{equation}
+	\label{eqn:eigenvalue_tail_bound_grid}
+	\lambda_{\mathbf{K}}(\wb{G}) \geq \frac{K^2}{N^2}
+	\end{equation}
+	for every $2 \leq K \leq N$.
+
+	\item 
+	\label{event:l2_to_empirical_norm_grid}
+	\textbf{Grid empirical norm:}
+	When $d \leq 4$, for every $f \in C^1(\Xset)$ satisfying $\norm{f}_{\Leb^2(\Xset)} \geq \sqrt{d/4} \cdot \norm{f}_{C^1(\Xset)} \cdot n^{-2/(4 + d)}$, additionally
+	\begin{equation}
+	\label{eqn:l2_to_empirical_norm_grid}
+	\frac{1}{n}\sum_{i = 1}^{n} \bigl(f(x_i)\bigr)^2 \geq \frac{1}{4}\norm{f}_{\Leb^2(\Xset)}^2
+	\end{equation}
+\end{enumerate}
+Once we have shown these inequalities, some basic algebra yields that for the specific choice $\kappa = (Ln)^{2d/(4 + d)}$ and a sufficiently large choice of $c$ in \eqref{eqn:holder_testing_rate_grid_1},
+\begin{align*}
+2b\sqrt{\frac{2\kappa}{n^2}} + \frac{f^T L f}{n\lambda_{\kappa}(\wb{G})} \leq \frac{1}{n}\sum_{i = 1}^{n}\bigl[f(x_i)\bigr]^2
+\end{align*}
+and~\eqref{eqn:holder_testing_rate_grid_2} follows from Lemma~\ref{lem:fixed_graph_testing}.
+
+\subsubsection{Proof of~\eqref{eqn:discrete_sobolev_norm_grid}.}
+The upper bound~\eqref{eqn:discrete_sobolev_norm_grid} follows straightforwardly from the Holder property,
+\begin{align*}
+f^T L f & = \sum_{i \in [N]^d} \sum_{j \in [N]^d} \bigl(f(\wb{x}_i) - f(\wb{x}_j\bigr)^2 \1\{\norm{i - j}_1 = 1\} \\
+& \leq \norm{f}_{C^1(\Xset)}^2 \sum_{i \in [N]^d} \sum_{j \in [N]^d} \norm{\wb{x}_i - \wb{x}_j}_2^2 \cdot  \1\{\norm{i - j}_1 = 1\} \\
+& = \norm{f}_{C^1(\Xset)}^2 n^{-2/d} \sum_{i \in [N]^d} \sum_{j \in [N]^d} \1\{\norm{i - j}_1 = 1\}  \\
+& \leq 2d \norm{f}_{C^1(\Xset)}^2 n^{1 - 2/d}.
+\end{align*}
+
+\subsubsection{Proof of~\eqref{eqn:eigenvalue_tail_bound_grid}.}
+The eigenvalues of $\wb{G}$ are known to have the analytic form
+\begin{equation*}
+\lambda_k(\wb{G}) = 4 \sum_{i = 1}^{d} \sin^2\Bigl(\frac{\pi(i_d - 1)}{2N}\Bigr)
+\end{equation*}
+whence the statement follows since $\sin(\pi x /2) \geq \min\{\pi x/2,\sqrt{2}/2 \}$ for all $x \in [0,1]$.
+
+\subsubsection{Proof of~\eqref{eqn:l2_to_empirical_norm_grid}.}
+For $x \in \Rd$, let $Q(x)$ be the $d$-dimensional cube of side length $1/N$ with $x$ at its center,
+\begin{equation*}
+Q(x) = [x_1 - 1/2N,x_1 + 1/2N] \otimes \cdots \otimes [x_d - 1/2N,x_d + 1/2N]
+\end{equation*}
+The difference between $f(\wb{x})$ and the average of $f$ over the cube $Q(\wb{x})$ can be bounded using the Holder property,
+\begin{equation*}
+\int_{Q(\wb{x})} (f(x) - f(\wb{x}))^2 \,dx \leq \mathrm{Leb}\Bigl(Q(\wb{x})\Bigr) \norm{f}_{C^1(\Xset)}^2 \frac{d}{4N^2} = \frac{d}{4n^{1 + 2/d}} \norm{f}_{C^1(\Xset)}^2
+\end{equation*}
+Dividing $\Xset$ into cubes and summing over the cubes, we have
+\begin{align*}
+\norm{f}_{\Leb^2(\Xset)}^2 & = \sum_{i \in [N]^d} \int_{Q(\wb{x}_i)} \bigl(f(x)\bigr)^2 \,dx \\
+& \leq 2 \sum_{i \in [N]^d} \Bigl\{\mathrm{Leb}\Bigl(Q(\wb{x})\Bigr) \bigl(f(\wb{x}_i)\bigr)^2 + \int_{Q(\wb{x}_i)} \bigl(f(x) - f(\wb{x}_i)\bigr)^2 \,dx \Bigr\} \\
+& \leq \frac{2}{n}\sum_{i \in [N]^d} \Bigl\{ \bigl(f(\wb{x}_i)\bigr)^2\Bigr\} + \frac{d}{4n^{2/d}} \norm{f}_{C^1(\Xset)}^2 \\
+& \leq \frac{2}{n}\sum_{i \in [N]^d} \Bigl\{ \bigl(f(\wb{x}_i)\bigr)^2\Bigr\} + \frac{1}{2}\norm{f}_{\Leb^2(\Xset)}^2
+\end{align*}
+where the last inequality follows by the assumption $\norm{f}_{\Leb^2(\Xset)}^2 \geq (d/4) \norm{f}_{C^1(\Xset)}^2 n^{-4/(4 + d)}$ along with the fact that $4/(4 + d) \leq 2/d$ when $d \leq 4$.
+
+\subsection{Proof of Theorem~\ref{thm:holder_testing_rate_grid_higher_order}}
+\label{subsec:holder_testing_rate_grid_ho_pf}
+We want to apply Lemma~\ref{lem:fixed_graph_testing} to the case where $G = \wb{G}$ and $s \geq 2$.
+In order to do this, we will need to show that when $\kappa$ is appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius} holds with respect to $G = \wb{G}$ and $\beta = (f(\wb{x}_i): i \in [N]^d)$.
+We already know  from~\eqref{eqn:eigenvalue_tail_bound_grid}, that
+\begin{equation*}
+\lambda_{\mathbf{K}}(\wb{G}) \geq \frac{K^2}{N^2}
+\end{equation*}
+for every $2 \leq K \leq N$. In addition, we will show that each of the following statements hold:
+\begin{enumerate}[label=(E\arabic*)]
+	\item
+	\textbf{Lattice Sobolev norm:}
+	\label{event:grid_sobolev_norm_higher_order}
+	There exists a constant $c$ such that for any $f \in C_c^s(\Xset)$,
+	\begin{equation}
+	\label{eqn:grid_sobolev_norm_higher_order}
+	f^T \wb{L}^s f \leq c \norm{f}_{C^s(\Xset)}^2 n^{1 - 2s/d}.
+	\end{equation}	
+	\item 
+	\textbf{Grid empirical norm:}
+	\label{event:grid_empirical_norm_higher_order}
+	There exists a sufficiently small constant $c_1$ and a sufficiently large constant $c_2$ such that for any $f \in C^s(\Xset;L)$,
+	\begin{equation}
+	\label{eqn:grid_empirical_norm_higher_order}
+	c_1 \norm{f}_n^2 \geq \norm{f}_{\Leb^2(\Xset)}^2 - c_2 L n^{-2s/d}
+	\end{equation}
+	for all $n$ sufficiently large.
+\end{enumerate}
+Then for the specific choice $\kappa = n^{2d/(4s + d)}$ and a sufficiently large choice of constant $c$ in \eqref{eqn:holder_testing_rate_grid_1},
+\begin{align*}
+2b\sqrt{\frac{2\kappa}{n^2}} + \frac{f^T \wb{L}^s f}{n\bigl(\lambda_{\kappa}(\wb{G})\bigr)^s} \leq \frac{1}{n}\sum_{i = 1}^{n}\bigl[f(x_i)\bigr]^2
+\end{align*}
+and~\eqref{eqn:holder_testing_rate_grid_2} follows from Lemma~\ref{lem:fixed_graph_testing}.
+
+\subsubsection{Proof of~\eqref{eqn:grid_sobolev_norm_higher_order}.}
+To prove~\eqref{eqn:grid_sobolev_norm_higher_order}, we approximate the action $\wb{L}f(\wb{x})$ to within a factor of $h^s$ for all $\wb{x} \in \wb{X}$.
+The manner in which we do this depends on whether $\wb{x}$ is sufficiently in the interior, or conversely is near the boundary, of $\wb{X}$.
+With an appropriate approximation in hand, we then find the desired estimate on the discrete semi-norm $f^T \wb{L}^s f$.
+
+\paragraph{Approximation at interior points.}
+Put $h = 1/N$, and for $j = 1,\ldots,d$, let $D_{e_j^2}f := f(\wb{x} + he_i) + f(\wb{x} - he_i) - 2f(\wb{x})$.
+At any point $\wb{x} \in \Xset_{h} \cap \wb{X} =: \wb{X}_{h}$, by taking a $2$nd order Taylor expansion of $f$ at $\wb{x}$ we can express the action of the lattice Laplacian as
+\begin{align}
+\wb{L}f(\wb{x}) & = \sum_{i = 1}^{d} D_{e_i^2}f(\wb{x}) \nonumber \\
+& = \sum_{i = 1}^{d} \biggl[\frac{\partial}{\partial e_i} f(\wb{x})(e_ih - e_ih) + \frac{h^2}{2} \int_{-h}^{h} \frac{\partial^2}{\partial e_i^2} f(\wb{x} + e_it) \,dt \biggr] \nonumber \\
+& = \frac{h^2}{2} \int_{-h}^{h} \sum_{i = 1}^{d} \frac{\partial^2}{\partial e_i^2} f(\wb{x} + e_it) \,dt \label{eqn:grid_sobolev_norm_higher_order_pf1}
+\end{align} 
+as a second-order differential operator which roughly approximates the continuum Laplace operator $\Delta f = -\sum_{i = 1}^{d} \frac{\partial^2}{\partial e_i^2} f$.
+Motivated by this, in Lemma~\ref{lem:grid_laplacian_approximation_error} we provide estimates on the absolute difference between the action of the iterated grid Laplacian and the iterated continuum Laplace operator.
+
+\begin{lemma}
+	\label{lem:grid_laplacian_approximation_error}
+	Let $f \in C^s(\Xset;L)$ for $s \geq 3$, and let $q = (s - 1)/2$ or $q = (s - 2)/2$. At all grid points $\wb{x} \in \wb{X}_{qh}$,
+	\begin{equation*}
+	\abs{\wb{L}^{q}f(\wb{x}) - h^{2q}\Delta^{q}f(\wb{x})} \leq cLh^s.
+	\end{equation*}
+\end{lemma} 
+We prove Lemma~\ref{lem:grid_laplacian_approximation_error} in Section~\ref{subsec:grid_laplacian_approximation_error_pf}.
+
+\paragraph{Approximation at boundary points.}
+For those remaining points $\wb{x} \in \wb{X} \cap \Xset_{qh}$ sufficiently near the boundary of $\Xset$, as might be expected to analyze $f(\wb{x})$ we will use the boundary assumptions we've made on $f$.
+Since $f$ is compactly supported in $\Xset$ there exists a point $x_0$ and some $\delta > 0$ such that $\mathrm{dist}(\wb{x},x_0) < qh$, and $f(z) = 0$ for all $z \in B(x_0,\delta)$.
+Therefore the partial derivatives $f^{(q)}(x_0) = 0$ for all $\abs{q} = 0,\ldots,s - 1$, and by taking a Taylor expansion of $f(x)$ around $f(x_0)$ we obtain
+\begin{equation*}
+\abs{f(x)} \leq q L h^{s}
+\end{equation*}
+Using this, we obtain the crude bound
+\begin{equation*}
+\abs{\wb{L}^s f(\wb{x})} \leq 2^s \cdot \max_{\wb{z} \in Q_{2qh}(\wb{x})} \bigl\{\abs{f(\wb{z})} \bigr\} \leq 2^s q L h^{s},
+\end{equation*}
+which will suffice for our purposes.
+
+\paragraph{Bounding the semi-norm.}
+When $s$ is even, our previous manipulations imply that
+\begin{align*}
+f^T \wb{L}^s f & \leq \sum_{\wb{x} \in \wb{X}_{qh}} \Bigl\{\sum_{i = 1}^d D_{e_i^2} \wb{L}^{(s - 2)/2}f(\wb{x}) \Bigr\}^2 + cL^2h^{2s}n^{1 - 1/d} \\
+& \leq h^{2s - 4}\sum_{\wb{x} \in \wb{X}_{qh}} \Bigl\{\sum_{i = 1}^d D_{e_i^2} \Delta^{(s - 2)/2}f(\wb{x}) \Bigr\}^2 + cL^2h^{2s}n^{1 - 1/d} \\
+& \leq \frac{h^{2s}}{4}\sum_{\wb{x} \in \wb{X}_{qh}} \Bigl\{\sum_{i = 1}^d \int_{-h}^{h} \Bigl(\frac{\partial^2}{\partial e_i^2}\Delta^{(s - 2)/2}f\Bigr)(\wb{x} + te_i) \,dt \Bigr\}^2 + cL^2h^{2s}n^{1 - 1/d} \\
+& \leq d^2 h^{2s} L^2 n + cL^2h^{2s}n^{1 - 1/d}
+\end{align*}
+where we have used, in order, our bounds on the absolute value of $\wb{L}^sf(\wb{x})$ and $\abs{f}(\wb{x})$ at points $\wb{x}$ near the boundary, our estimate on the approximation error between the grid Laplacian and continuum Laplace operator,
+the representation~\eqref{eqn:grid_sobolev_norm_higher_order_pf1}, and finally the Holder property
+\begin{equation*}
+\norm{\Delta^{(s - 2)/2}f}_{C^2(\Xset)} \leq \norm{f}_{C^s(\Xset)} \leq L.
+\end{equation*}
+Similar manipulations supply an equivalent bound when $s$ is odd
+\begin{align*}
+f^T \wb{L}^s f & \leq \sum_{\wb{x} \in \wb{X}_{qh}} \Bigl\{\sum_{i = 1}^d \wb{L}^{(s - 1)/2}f(\wb{x} + he_i) - \wb{L}^{(s - 1)/2}f(\wb{x}) \Bigr\}^2 + cL^2h^{2s}n^{1 - 1/d} \\
+& \leq h^{2s - 2}\sum_{\wb{x} \in \wb{X}_{qh}} \Bigl\{\sum_{i = 1}^d \wb{\Delta}^{(s - 1)/2}f(\wb{x} + he_i) - \wb{\Delta}^{(s - 1)/2}f(\wb{x}) \Bigr\}^2 + cL^2h^{2s}n^{1 - 1/d} \\
+& \leq d^2 h^{2s} L^2 n + cL^2h^{2s}n^{1 - 1/d}
+\end{align*}
+where this time we have invoked the Holder property $\norm{\Delta^{(s - 1)/2}f}_{C^1(\Xset)} \leq \norm{f}_{C^s(\Xset)} \leq L$. This establishes~\eqref{eqn:grid_sobolev_norm_higher_order}.
+
+\subsubsection{Proof of~\eqref{eqn:grid_empirical_norm_higher_order}.}
+\label{subsubsec:grid_empirical_norm_ho_pf}
+Take $\eta$ to be the tensorization of a univariate $(s - 1)$-order kernel $\nu$. More specifically, we let $\nu: \Reals \to \Reals$ satisfy
+\begin{equation*}
+\int_{\Reals} \nu(x) \,dx = 1,~~ \int x^{\ell} \nu(x) \,dx = 0~~\textrm{for $\ell = 1,\ldots,s-1$},~~ \abs{\nu(x)} \leq 1~\textrm{for all $x \in [0,1^d]$}
+\end{equation*}
+additionally be compactly supported on $[-1/2,1/2]$ but otherwise arbitrary, and define $\eta:\Reals^d \to \Reals$ to be $\eta(z) = \prod_{i = 1}^{d} \nu(x_i)$ for $z \in \Reals^d$.
+We denote $\eta_h(x) = \frac{1}{h^d} \eta(x/h)$, and note that for any polynomial $u$ of degree at most $s - 1$, $\bigl(u \ast \eta_h\bigr)(x) = u(x)$.
+
+Fix an odd integer $m \geq 1$ to be chosen large enough, but fixed, later on, and assume without loss of generality that $N/m =: M \in \mathbb{N}$.
+For $i \in [n]^d$, let $u_{i}$ be the $(s - 1)$-order Taylor expansion of $f$ around $\wb{x}_i$, and further let
+\begin{align*}
+u(x) = \sum_{i \in [M]^d} u_{mi - 1}(x) \cdot \1\{x \in Q_{hm}(\wb{x}_{mi - 1})  \},~~ \wt{u}(x) = \sum_{i \in [M]^d} \sum_{\wb{x} \in  Q_{hm}(\wb{x}_{mi - 1})} \bigl(u_{mi - 1} \ast \eta_h\bigr)(\wb{x}) \cdot \1\{x \in Q_{h}(\wb{x})  \}.
+\end{align*}
+be piecewise polynomial and piecewise constant functions, respectively.
+
+Standard properties of Taylor expansions imply that $\abs{f(x) - u_{mi - 1}(\wb{x})} \leq (L\sqrt{d}mh)^{s}$ for all $x \in Q_{mh}(\wb{x}_{mi - 1})$, and as a result
+\begin{equation*}
+\norm{u - f}_{\Leb^2([0,1]^d)}^2,~~\norm{u - f}_{n}^2 \leq ch^{2s}
+\end{equation*}
+Since $u_{mi- 1}$ is an $(s - 1)$-order polynomial, we have that $\bigl(u_{mi - 1} \ast \eta_h\bigr)(\wb{x}) = u_{mi - 1}(\wb{x})$ for each $\wb{x} \in \wb{X}$; therefore $\norm{\wt{u}}_n = \norm{u}_n$.
+Additionally $\wt{u}$ is piecewise constant over the grid cells $Q_h(\wb{x})$, and so we have
+\begin{equation*}
+\norm{\wt{u}}_{\Leb^2([0,1]^d)}^2 = \sum_{\wb{x} \in \wb{X}} h^d \cdot  [\wt{u}(\wb{x})]^2 = \norm{\wt{u}}_n^2.
+\end{equation*}
+It remains only to show that
+\begin{equation*}
+c \norm{u}_{\Leb^2([0,1]^d)} \leq \norm{\wt{u}}_{\Leb^2([0,1]^d)},
+\end{equation*}
+for a constant $c$ which does not depend on $\wt{u}$.
+In Lemma~\ref{lem:riemann_sums_polynomials}, we show that the statement holds true if $u$ is a polynomial on $[0,1]^d$; and the width of the partition used to form $\wt{u}$ is taken to be a sufficiently small constant.
+By translating and rescaling the cubes $Q_{mi - 1}(\wb{x}_{mi} - 1)$ into $[0,1]^d$ as in \textcolor{red}{(Arias-Castro)}, we have that the same result holds for the piecewise polynomial $u$ when $m$ is taken to be a sufficiently large constant.
+Putting the pieces together, we have that whenever $n$ is great enough such that $M \in \mathbb{N}$,
+\begin{align*}
+\norm{f}_{\Leb^2([0,1]^d)}^2 & \leq cn^{-2s/d} + 2\norm{u}_{\Leb^2([0,1]^d)}^2 \\ & \leq cn^{-2s/d} + c\norm{\wt{u}}_{\Leb^2([0,1]^d)}^2 \\
+& = cn^{-2s/d} + c\norm{u}_{n}^2 \\
+& \leq cn^{-2s/d} + c\norm{f}_n^2.
+\end{align*}
+which completes the proof of~\eqref{eqn:grid_empirical_norm_higher_order}.
+
+\subsection{Proof of Theorem~\ref{thm:sobolev_testing_rate_grid}}
+
+We want to apply Lemma~\ref{lem:fixed_graph_testing} to the case where $G = \wb{G}$.
+In order to do this, we will need to show that when $\kappa$ is appropriately tuned and $\norm{f}_{\Leb^2(\mathcal{X})}$ is sufficiently large, the inequality~\eqref{eqn:fixed_graph_testing_critical_radius_1} holds with respect to $G = \wb{G}$ and $\beta = (f(\wb{x}_i): i \in [N]^d)$.
+We will show that for all $f = \sum_{k \in \Nbb^d} \theta_k \varphi_k \in \wt{H}^s(\Xset;L)$,
+\begin{equation}
+\label{eqn:grid_sobolev_continuum_sobolev_approximation}
+\norm{\Pi_{\kappa,G}(f)}_n^2 \geq \norm{f}_{\Leb^2}^2 - c L \norm{f}_{\Leb^2(\Xset)} \cdot n^{-s/d} - \sum_{k \in \mathbb{N}^d/[K]^d} \theta_k^2 
+\end{equation}
+The tail sum of $\theta$ on the right hand side of~\eqref{eqn:grid_sobolev_continuum_sobolev_approximation} can be upper bounded,
+\begin{equation*}
+ \sum_{k \in \mathbb{N}^d/[K]^d} \theta_k^2  \leq L^2 K^{-2s} = L^2 \kappa^{-2s/d}
+\end{equation*}
+and since $2s > d$, we have that for sufficiently large $n$
+\begin{equation*}
+c L \norm{f}_{\Leb^2(\Xset)} \cdot n^{-s/d} \leq \frac{1}{2}\norm{f}_{\Leb^2(\Xset)}^2.
+\end{equation*}
+Along with~\eqref{eqn:grid_sobolev_continuum_sobolev_approximation} this implies
+\begin{equation*}
+\norm{\Pi_{\kappa,G}(f)}_n^2 \geq \frac{1}{2} \norm{f}_{\Leb^2}^2 - L^2 \kappa^{-2s/d}.
+\end{equation*}
+Then for the specific choice $\kappa = n^{2d/(4s + d)}$ and a sufficiently large choice of constant $c$ in~\eqref{eqn:sobolev_testing_rate_grid_1}, 
+\begin{equation*}
+\norm{f}_{\Leb^2}^2 \geq L^2 \kappa^{-2s/d} + 2b\sqrt{\frac{2\kappa}{n^2}}
+\end{equation*}
+which establishes~\eqref{eqn:fixed_graph_testing_critical_radius_1}, allowing us to apply Lemma~\ref{lem:fixed_graph_testing} and completing the proof of Theorem~\ref{thm:sobolev_testing_rate_grid}.
+
+\subsubsection{Proof of~\eqref{eqn:grid_sobolev_continuum_sobolev_approximation}.}
+
+Recall that the eigenvectors of the grid graph $v_k := v_k(\wb{G}_d)$ satisfy
+\begin{equation*}
+v_{k}(\wb{x}_i) = \frac{1}{\sqrt{n}} \varphi_{k}(\wb{x}_i)
+\end{equation*}
+for each $i \in [N]^d$ and $k \in \mathbb{N}^d$. Therefore we may rewrite $\Pi_{\kappa,\wb{G}_d}$ as a function of the tensor product cosine Fourier basis,
+\begin{align*}
+\Pi_{\kappa,\wb{G}_d}(f) & = \frac{1}{n} \sum_{k \in [K]^d} \Bigl\{\sum_{i \in [N]^d} f(\wb{x}_i) \varphi_k(\wb{x}_i)\Bigr\} \varphi_k \\
+& :=  \sum_{k \in [K]^d} \wt{\theta}_k \varphi_k
+\end{align*}
+where we interpret the right hand side as a function over $\wb{X}$. Since $(\varphi_k)$ is an $L_2(\wb{X})$ orthonormal sequence, we obtain
+\begin{align*}
+\norm{\Pi_{\kappa,\wb{G}_d}(f)}_n^2 & = \sum_{k \in [K]^d} \wt{\theta}_k^2 \\ 
+& = \sum_{k \in [K]^d} {\theta}_k^2 + \sum_{k \in [K]^d} \wt{\theta}_k^2 - \theta_k^2 \\
+& = \norm{f}_{\Leb^2([0,1]^d)}^2 - \sum_{k \in \Nbb^d \setminus [K]^d} \theta_k^2 + \sum_{k \in [K]^d} \wt{\theta}_k^2 - \theta_k^2
+\end{align*}
+or equivalently
+\begin{equation*}
+\norm{f}_{\Leb^2([0,1]^d)}^2 - \norm{\Pi_{\kappa,\wb{G}_d}(f)}_n^2 = \sum_{k \in [K]^d} \theta_k^2 - \wt{\theta}_k^2 + \sum_{k \in \Nbb^d \setminus [K]^d} \theta_k^2
+\end{equation*}
+The second term in the above equation is exactly the third term in~\eqref{eqn:grid_sobolev_continuum_sobolev_approximation}, and so we focus our attention on the first term, which represents discretization error.
+Applying the parallelogram law and the Cauchy-Schwarz inequality gives
+\begin{equation}
+\label{eqn:grid_sobolev_approximation_error6_pf1}
+\begin{aligned}
+\sum_{k = 1}^{\kappa} \theta_k^2 - \wt{\theta}_k^2 & \leq \sum_{k \in [K]^d} \max\Bigl\{(\theta_k + \wt{\theta}_k)(\theta_k - \wt{\theta}_k),0\Bigr\} \\
+& \leq \left(\sum_{k \in [K]^d} \max\Bigl\{(\theta_k + \wt{\theta}_k)^2,0\Bigr\}\right)^{1/2} \left(\sum_{k \in [K]^d} (\theta_k - \wt{\theta}_k)^2\right)^{1/2} \\
+& \leq \sqrt{2} \norm{f}_{\Leb^2([0,1])} \left(\sum_{k \in [K]^d} (\theta_k - \wt{\theta}_k)^2\right)^{1/2}
+\end{aligned}
+\end{equation}
+
+It is well known that the error $\theta_k - \wt{\theta}_k$ is due to the aliasing phenomenon, where high frequency sinusoidal waves agree with low frequency waves at evenly spaced points.
+This phenomenon has a very precise periodic nature---which we formally state in Lemma~\ref{lem:alias}---which allows us to rewrite $\wt{\theta}_k$.
+Note that for every $\ell \in \mathbb{N}^d$, there exists exactly one $m \in \mathbb{N}_0^d$, $o \in [N]^d$ and $b \in \{0,1\}^d$ such that $\ell = 2mN + (1 - b)o - b((o - 1))$ (where for $a,b \in \Reals^d$ we denote $ab := (a_1b_1,\ldots,a_db_d)$).
+Let
+\begin{equation*}
+I_N(m,b,o) = 
+\begin{cases*}
+2mn + k,& ~~\textrm{if $b = 0, k \in [N]$,} \\
+2mn - (k - 2),& ~~\textrm{if $b = 1, k \in [N], k > 1$}, \\
+0,& ~~\textrm{if $b = 1$ and $k = 1$.}
+\end{cases*}
+\end{equation*}
+Adopting the convention $\theta_{\ell} := 0$ if $\ell$ contains a non-positive index, we have
+\begin{align*}
+\wt{\theta}_k & = \frac{1}{n} \sum_{i \in [N]^d} f(\wb{x}_i) \varphi_k(\wb{x}_i) \\
+& = \frac{1}{n} \sum_{i \in [N]^d} \Bigl(\sum_{\ell \in \Nbb^d} \theta_{\ell} \varphi_{\ell}(\wb{x}_i) \Bigr) \varphi_k(\wb{x}_i) \\
+& = \sum_{\ell \in \mathbb{N}^d} \theta_{\ell} \Bigl(\frac{1}{n} \sum_{i \in [N]^d} \varphi_{\ell}(\wb{x}_i) \varphi_{k}(\wb{x}_i)\Bigr) \\
+& = \sum_{m \in \mathbb{N}_0^d} \sum_{o \in [N]^d}\sum_{b \in \{0,1\}^d} \theta_{I_N(m,b,o)}  \Bigl(\frac{1}{n} \sum_{i \in [N]^d} \varphi_{I_N(m,b,o)}(\wb{x}_i) \varphi_{k}(\wb{x}_i)\Bigr).
+\end{align*}
+We now derive an upper bound on the magnitude of the error $|\theta_k - \wt{\theta}_k|$ by using first Lemma~\ref{lem:alias}, second the Cauchy-Schwarz inequality, and third the fact that whenever $I_N(m,b,k) > 0$, we have $I_N(m,b,k) \geq mN$ (where inequality between two vectors is interpreted entrywise).
+\begin{align*}
+\abs{\wt{\theta}_k - \theta_k} & \leq 2^{d/2} \sum_{m \in \Nbb_0^d \setminus \{0\}} \sum_{b \in \{0,1\}^d} \abs{\theta_{I_N(m,b,k)}} \\
+& \leq 2^{d/2} \biggl(\sum_{m \in \Nbb_0^d \setminus \{0\}} \sum_{b \in \{0,1\}^d} \Bigl(\theta_{I_N(m,b,k)}a_{I_N(m,b,k)}\Bigr)^2 \biggr)^{1/2} \biggl(\sum_{m \in \Nbb_0^d \setminus \{0\}} \sum_{b \in \{0,1\}^d} \Bigl(a_{I_N(m,b,k)}\Bigr)^{-2} \biggr)^{1/2} \\
+& \leq 2^{d}N^{-s } \biggl(\sum_{m \in \Nbb_0^d \setminus \{0\}} \sum_{b \in \{0,1\}^d} \Bigl(\theta_{I_N(m,b,k)}a_{I_N(m,b,k)}\Bigr)^2 \biggr)^{1/2} \biggl(\sum_{m \in \Nbb_0^d \setminus \{0\}}\sum_{j = 1}^{d} m_j^{-2s} \biggr)^{1/2}
+\end{align*}
+Letting $c(d,s)$ be a finite constant which may change from line to line but can depend only on $d$ and $s$, and recalling the notation $\Nbb_2 = \Nbb_0 \setminus \{0,1\}$, we can upper bound the second sum in the previous display,
+\begin{align*}
+\sum_{m \in \Nbb_0^d \setminus \{0\}}\sum_{j = 1}^{d} m_j^{-2s} & \overset{(i)}{\leq} d^{2(s + 1)} \sum_{m \in \Nbb_0^d \setminus \{0\}} \Bigl(\sum_{j = 1}^{d} m_j^{-2}\Bigr)^s \\
+& \overset{(ii)}{\leq} d^{2(s + 1)} \sum_{\wt{d} = 1}^{d} 2^{\wt{d}} \frac{d!}{\wt{d}!} \sum_{m \in \Nbb_2^{\wt{d}}} \Bigl(\sum_{j = 1}^{d} m_j^{-2}\Bigr)^s \\
+& \overset{(iii)}{\leq} d^{2(s + 1)} \sum_{\wt{d} = 1}^{d} 2^{\wt{d}} \frac{d!}{\wt{d}!} \int_{[1,\infty)^{\wt{d}}} \Bigl(\sum_{j = 1}^{\wt{d}} x_j^{-2}\Bigr)^{s} \,dx_1 \ldots \,dx_{\wt{d}} \\
+& \leq \sum_{\wt{d} = 1}^{d} c(d,s) \int_{\sqrt{\wt{d}}}^{\infty} r^{-2s + d - 1} \,dr \\
+& \overset{(iv)}{\leq} c(d,s) < \infty.
+\end{align*}
+where $(i)$ is an equality when $d = 1$, and otherwise follows from Jensen's inequality when $d \geq 2$ and $s \geq 1$; $(ii)$ follows from specially considering the edge cases where $m_j = 0$ or $m_j = 1$ for $j = \wt{d} + 1,\ldots, d$; $(iii)$ uses the fact that the Riemann sum of a monotone non-increasing function evaluated at right end points is always no greater than the corresponding integral; and $(iv)$ follows since $2s > d$.
+We are now in a position to conclude that
+\begin{align}
+\sum_{k \in [K]^d} (\theta_k - \wt{\theta}_k)^2 & \leq c(d,s) N^{-2s} \sum_{k \in [K]^d} \sum_{m \in \Nbb_0^d \setminus \{0\}} \sum_{b \in \{0,1\}^d} \Bigl(\theta_{I_N(m,b,k)}a_{I_N(m,b,k)}\Bigr)^2 \nonumber \\
+& \leq c(d,s) N^{-2s} \sum_{k \in \Nbb^d \setminus [N]^d} \theta_k^2 a_k^2 \nonumber \\
+& \leq L c(d,s) N^{-2s} \label{eqn:grid_sobolev_approximation_error6_pf2}
+\end{align}
+where the second inequality in the preceding display follows since for any $(m,k,b) \neq (m',k',b')$, either $I_N(m,b,k) \neq I_N(m',b',k')$ or $I_N(m,b,k) = I_N(m',b',k') = 0$.
+We plug~\eqref{eqn:grid_sobolev_approximation_error6_pf2} back in to~\eqref{eqn:grid_sobolev_approximation_error6_pf1} to obtain~\eqref{eqn:grid_sobolev_continuum_sobolev_approximation}.
+
+\subsection{Proof of Lemma~\ref{lem:holder_testing_rate_grid_low_smoothness_lb}}
+We note that when $d = 1$, for all $i \in [n]$
+\begin{equation*}
+\varphi_{n + 1}(\wb{x}_i) = \sqrt{2}\cos\Bigl(\frac{\pi}{2}(2i - 1)\Bigr) = 0.
+\end{equation*}
+and therefore for $\mathbf{M} = (N + 1,\ldots,N + 1)$ and all $i \in [N]^d$,
+\begin{equation*}
+\varphi_{\mathbf{M}}(\wb{x}_i) = 0.
+\end{equation*}
+Letting 
+\begin{equation*}
+a = \frac{L}{2^{d/2}\pi^s(d+1)^s}N^{-s},~~ g = a \varphi_{N+1}
+\end{equation*}
+some simple calculations show that $g$ satisfies the conditions of Lemma~\ref{lem:holder_testing_rate_grid_low_smoothness_lb}.
+
+\subsection{Proof of Proposition~\ref{prop:holder_testing_rate_grid_low_smoothness_ub}}
+Let $f \in C^s(\Xset;L)$ satisfy
+\begin{equation*}
+\norm{f}_{\Leb}^2 \geq c b L^2 n^{-2s/d}
+\end{equation*}
+for a sufficiently large constant $c$.
+Note that when $\kappa = n$, $\pi_{\kappa,\wb{G}}(f) = f$ when the latter is considered as a function on $\wb{X}$, i.e.
+a vector.
+Therefore the claimed result follows from Lemma~\ref{lem:fixed_graph_testing}, once we show
+\begin{equation*}
+\frac{1}{n} \sum_{i \in [N]^d} \bigl(f(\wb{x}_i)\bigr)^2 \geq 2b \sqrt{\frac{2}{n}}
+\end{equation*}
+However, it follows from~\eqref{eqn:grid_empirical_norm_higher_order} that the above inequality is satisfied whenever
+\begin{equation*}
+\frac{1}{n} \sum_{i \in [N]^d} \bigl(f(\wb{x}_i)\bigr)^2 \geq \max\{c L^2 n^{-2s/d}, b n^{-1/2}\} \leq c b L^2 n^{-2s/d}
+\end{equation*}
+when $4s \leq d$.
+
+\clearpage
 
@@
 & \leq c r^{2s} \norm{f}_{H^s(\Xset)}^2
-\label{eqn:expected_difference_operators_sobolev_pf1}
 \end{align*}
 where the last inequality follows by Lemma~\ref{lem:expected_first_order_seminorm}.
+
+\subsection{Proof of Lemma~\ref{lem:grid_laplacian_approximation_error}}
+\label{subsec:grid_laplacian_approximation_error_pf}
+We prove by induction on $q$.
+In the base case, $q = 1$, $s = 3$ or $s = 4$, and we assume $\wb{x} \in \wb{X}_{h}$.
+By taking Taylor expansions of $f(\wb{x} + e_ih)$ and $f(\wb{x} - e_ih)$ around $f(\wb{x})$ for $i = 1,\ldots,d$, we obtain
+\begin{align}
+\wb{L}f(\wb{x}) & = \sum_{i = 1}^{d} D_{e_i^2}f(\wb{x}) \nonumber \\
+& = \sum_{i = 1}^{d} \Biggl[\sum_{\alpha = 1}^{s - 1}\frac{1}{\alpha!}\frac{\partial^{\alpha}}{\partial e_i^{\alpha}} f(\wb{x})\bigl(h^{\alpha} + (-h)^{\alpha}\bigr) + R(\wb{x}) \Biggr] \nonumber \\
+& = \sum_{i = 1}^{d} \Biggl[\sum_{\beta = 1}^{\floor{(s - 1)/2}}\frac{2}{(2\beta)!}\frac{\partial^{2\beta}}{\partial e_i^{2\beta}} f(\wb{x}) h^{2\beta} + R(\wb{x})\Biggr] \label{eqn:grid_laplacian_approximation_error_pf1}
+\end{align} 
+where the remainder term is upper bounded
+\begin{equation*}
+\abs{R(\wb{x})} = \abs{\frac{1}{s!}\sum_{i = 1}^d \int_{-h}^{h} \frac{\partial^s}{\partial e_i^s} f(\wb{x} + e_it) t^{s - 1} \,dt} \leq \frac{d}{(s + 1)!} L h^s 
+\end{equation*}
+When $s = 3$ or $s = 4$, the first sum includes only second-order derivatives,
+\begin{equation*}
+\sum_{i = 1}^{d} \sum_{\beta = 1}^{\floor{(s - 1)/2}}\frac{2}{(2\beta)!}\frac{\partial^{2\beta}}{\partial e_i^{2\beta}} f(\wb{x}) h^{2\beta} = h^2 \sum_{i = 1}^d \frac{\partial^{2\beta}}{\partial e_i^{2\beta}} f(\wb{x}) = h^2 \Delta f(\wb{x}),
+\end{equation*}
+and the claim is shown.
+
+When $q \geq 2$, we use the prior analysis to rewrite the $q$th-order iterated Laplacian
+\begin{align}
+\wb{L}^qf(\wb{x}) & = \wb{L}^{q - 1} \Biggl(\sum_{i = 1}^d \sum_{\beta = 1}^{\floor{(s - 1)/2}} \frac{2}{(2\beta)!} \frac{\partial^{2\beta}}{\partial e_i^{2\beta}} f h^{2 \beta} + R\Biggr)(\wb{x}) \nonumber \\
+& = \sum_{\beta = 1}^{\floor{(s - 1)/2}} \frac{2h^{2\beta}}{(2\beta)!} \wb{L}^{q - 1} \biggl(\sum_{i = 1}^d\frac{\partial^{2\beta}}{\partial e_i^{2\beta}} f \biggr)(\wb{x}) + \wb{L}^{q - 1}R(\wb{x}) \label{eqn:grid_laplacian_approximation_error_pf5}
+\end{align}
+We deal with the remainder term first. Noting that $Q_{2(q - 1)h}(\wb{x}) \subset \Xset_{h}$, we have
+\begin{align}
+\abs{\wb{L}^{q - 1}R(\wb{x})} & \leq \bigl(2(q - 1)d\bigr)^{q - 1} \cdot \max_{\wb{z} \in Q_{2(q - 1)h}(\wb{x})} \bigl\{\abs{R(\wb{z})}\bigr\} \nonumber \\
+& \leq \frac{\bigl(2(q - 1)d\bigr)^{q - 1}d}{(s + 1)!} L h^s \label{eqn:grid_laplacian_approximation_error_pf2}
+\end{align}
+Next we deal with the higher-order terms in the first sum, meaning the terms where $\beta \geq 2$.
+For convenience, denote $f_\beta := \sum_{i = 1}^d\frac{\partial^{2\beta}}{\partial e_i^{2\beta}} f$, and note that $f_{\beta} \in C^{s - 2\beta}(\Xset;L)$.
+We will show that for any $\wb{x} \in \wb{X}_{qh}$,
+\begin{equation}
+\label{eqn:grid_laplacian_approximation_error_pf3}
+\abs{\wb{L}^{q - 1}f_{\beta}(\wb{x})} \leq c L h^{s - 2\beta}.
+\end{equation}
+which will be sufficient for the purposes of proving the Lemma.
+
+Let us assume $s - 2\beta$ is odd. Set $s' = (s - 2\beta - 1)/2$, and $s'' = q - 1 - (s' + 1)$, and note that $s''$ is a non-negative integer. By hypothesis, for any $\wb{z} \in \wb{X}_{hs'}$
+\begin{equation*}
+\abs{\wb{L}^{s'}f_{\beta}(\wb{z}) - h^{2s'}\Delta^{s'}f_{\beta}(\wb{z})} \leq cLh^{s - 2\beta}
+\end{equation*}
+and consequently,
+\begin{align*}
+\abs{\wb{L}^{q - 1}f_{\beta}(\wb{x})} & \leq (2s''d)^{s''} \cdot \max_{\wb{z} \in Q_{2s''h}(\wb{x})} \Bigl\{\abs{\wb{L}^{s' + 1}f(\wb{z})}\Bigr\} \\
+& \leq (2s''d)^{s''} \cdot \max_{\wb{z} \in \wb{X}_{(s' + 1)h}} \Bigl\{\abs{\wb{L}^{s' + 1}f(\wb{z})}\Bigr\}  \\
+& \leq (2s''d)^{s''} \Bigl( h^{2s} \cdot \max_{\wb{z} \in \wb{X}_{(s' + 1)h}} \Bigl\{ \abs{\wb{L} \Delta^{s'} f_{\beta}(\wb{z})} \Bigr\} + cLh^{s - 2\beta}\Bigr) \\
+& \leq (2s''d)^{s''} \Bigl( 2dL h^{2s' + 1}  + cLh^{s - 2\beta}\Bigr)
+\end{align*}
+where the last line follows since $\Delta^{s'}{f_{\beta}} \in C^1(\ol{X};L)$.
+Since $2s' + 1 = s - 2\beta$, we have shown the desired result~\eqref{eqn:grid_laplacian_approximation_error_pf3}.
+Similar manipulations prove the result when $s - 2\beta$ is even.
+
+Finally, we consider the second order term, where $\beta = 1$ and $f_\beta = \Delta f \in C^{s - 2}(\Xset;L)$.
+Then, since either $q - 1 = (s - 3)/2$ or $q - 1 = (s - 4)/2$, by hypothesis for any $\wb{x} \in \wb{X}_{qh} \subset \wb{X}_{(q - 1)h} $,
+\begin{equation}
+\label{eqn:grid_laplacian_approximation_error_pf4}
+h^{2}\abs{ \wb{L}^{q - 1} \Delta f(\wb{x}) - h^{2(q - 1)} \Delta^{q} f(\wb{x})  } \leq cL h^{2}h^{s - 2} = cLh^s
+\end{equation}
+The claimed result follows upon plugging~\eqref{eqn:grid_laplacian_approximation_error_pf2},\eqref{eqn:grid_laplacian_approximation_error_pf3}, and \eqref{eqn:grid_laplacian_approximation_error_pf4} back into~\eqref{eqn:grid_laplacian_approximation_error_pf5}.
 
@@
 
+To simplify the statement of the following Lemma---which we use in the proof of~\eqref{eqn:grid_empirical_norm_higher_order}---for $j \in \mathbb{N}$ let
+\begin{equation*}
+\mathcal{K}_j(u) := \sum_{[j]^d} \bigl(u \ast \eta_h\bigr)(\wb{x}) \cdot \1\{x \in Q_{1/j}(\wb{x})\}.
+\end{equation*}
+where $\eta_h$ is taken as in Section~\ref{subsubsec:grid_empirical_norm_ho_pf}. Denote the class of polynomials on $\Reals^d$ of degree at most $m$ by $\mathcal{P}_m^d$.
+
+\begin{lemma}
+	\label{lem:riemann_sums_polynomials}
+	There exists a constant $c$ such that for any $j > 1/c$, 
+	\begin{equation*}
+	\norm{\mathcal{K}_j(u)}_{\Leb^2([0,1]^d)} \geq c \norm{u}_{\Leb^2([0,1]^d)}
+	\end{equation*}
+	for all polynomials $u \in \mathcal{P}_{s - 1}^d$.
+\end{lemma}
+\begin{proof}
+	We proceed by contradiction, following the proof of~\textcolor{red}{(Arias-Castro)} closely and making necessary adjustments.
+Suppose there exists a sequence $(u_j) \subset \mathcal{P}_{s - 1}^d$ such that $\norm{\mathcal{K}_j(u)}_{\Leb^2([0,1]^d)} \leq 1/j \norm{u_j}_{\Leb^2([0,1]^d)}$ for each $j \in \mathcal{N}$.
+Assume without loss of generality that $\norm{u_j}_{\Leb^2([0,1]^d)} = 1$; then there exists an accumulation point $u_{\infty}$ of $(u_j)$, and $\norm{u_{\infty}}_{\Leb^2([0,1]^d)} = 1$.
+However, by Jensen's inequality, and the boundedness and compact support of $\eta$, we have
+	\begin{align*}
+	\Bigl[\bigl(u \ast \eta_h\bigr)(\wb{x})\Bigr]^2 & = \Bigl[\int u(y) \eta_h(y - \wb{x}) \,dy \Bigr]^2 \\
+	& = \biggl[\frac{1}{h^d}\int u(y) \eta\Bigl(\frac{y - \wb{x}}{h}\Bigr) \,dy \biggr]^2 \\
+	& \leq \frac{1}{h^d}  \biggl[u(y) \eta\Bigl(\frac{y - \wb{x}}{h}\Bigr)\biggr]^2 \,dy \\
+	& \leq \frac{1}{h^d} \int_{Q_h(\wb{x})} \bigl[u(y)\bigr]^2 \,dy
+	\end{align*}
+	which in turn gives 
+	\begin{align*}
+	\norm{\mathcal{K}_j(u)}_{\Leb^2([0,1]^d)}^2 & \leq \sum_{\wb{x} \in \wb{X}} h^d\Bigl[\bigl(u \ast \eta_h\bigr)(\wb{x})\Bigr]^2 \\
+	& \leq \sum_{\wb{x} \in \wb{X}}\int_{Q_h(\wb{x})} \bigl[u(y)\bigr]^2 \\
+	& = \norm{u}_{\Leb^2([0,1]^d)}^2.
+	\end{align*}
+	As a result,
+	\begin{align*}
+	\norm{\mathcal{K}_j(u_{\infty})}_{\Leb^2([0,1]^d)} & \leq \norm{\mathcal{K}_j(u_{\infty} - u_j)}_{\Leb^2([0,1]^d)} + \norm{\mathcal{K}_j(u_j)}_{\Leb^2([0,1]^d)} \\
+	& \leq \norm{u_\infty - u_j}_{\Leb^2([0,1]^d)} + 1/j \to 0
+	\end{align*}
+	as $j \to \infty$.
+On the other hand, $\mathcal{K}_j(u_{\infty}) \to u_{\infty}$ as $j \to \infty$.
+In light of the boundedness of $\eta$ and the boundedness of $u_{\infty}$ on $[0,1]^d$, we may apply the dominated convergence theorem and obtain that $\norm{\mathcal{K}_j(u_{\infty})}_{\Leb^2([0,1]^d)} \to \norm{u_{\infty}}_{\Leb^2([0,1]^d)}$.
+This establishes a contradiction, thus proving the claim.
+\end{proof}
+
 \subsection{One-Sided Concentration}
@@
 
+\subsection{Discrete norms}
+\begin{lemma}
+	\label{lem:alias}
+	Let $k \in [n]$ and let $\ell \in \Nbb$. There exists a unique $m \in \Nbb_{0}$, $j \in [n]$ and $b \in \{0,1\}$ such that $\ell = 2mn + (1 - b)j - b(j - 1)$. Then
+	\begin{equation*}
+	\abs{\frac{1}{n}\sum_{i = 1}^{n} \varphi_{k}(\wb{x}_i)\varphi_{\ell}(\wb{x}_i)} \leq  
+	\begin{cases*}
+	2^{d/2},& ~~ \textrm{if $b = 0$ and $j = k$, or $b = 1$ and $j = k - 1$} \\
+	0,& ~~ \textrm{otherwise.}
+	\end{cases*}
+	\end{equation*}
+\end{lemma}
+
+Note that it is possible to sharpen the factor $2^{d/2}$ to $1$ whenever $k \neq 1$.
+Since we are not too concerned with constants which depend only $d$, we simply lump everything together into the weaker bound.
+
+\begin{proof}
+	Note that for $k = 1,\ldots,n$, and $i = 1,\ldots,n$,
+	\begin{equation*}
+	\sqrt{n}v_{k,i}(\wb{G}_1) = \varphi_k(\wb{x}_i).
+	\end{equation*}
+	Additionally, it is easy to check that
+	\begin{equation*}
+	\varphi_{\ell}(\wb{x}_i) \propto
+	\begin{cases*}
+	\varphi_{j}(\wb{x}_i),& ~~\textrm{if $b = 0$} \\
+	\varphi_{j + 1}(\wb{x}_i),& ~~\textrm{if $b = 1$.}
+	\end{cases*}
+	\end{equation*}
+	where the constant of proportionality is always no greater than $2^{d/2}$.
+	The claim follows since the eigenvectors $\sqrt{n} v_k(\wb{G}_1)$ are orthonormal in empirical norm.	
+\end{proof}
+
+\subsection{Equivalence between Sobolev classes}
+\label{subsec:sobolev_class_equivalence}
+
+We will show that when $s \geq 1$,
+\begin{equation*}
+\wt{H}^s([0,1];L) = \Bigl\{f \in H^{s}([0,1];L'): f^{(\ell)}(0) = f^{(\ell)}(1) = 0,~~\textrm{for each $\ell < s$ odd.} \Bigr \}
+\end{equation*}
+First, we show the $\subseteq$ direction.
+Let $f = \sum_{k = 1}^{\infty} \theta_k \varphi_k$ satisfy $\sum_{k = 1}^{\infty} \theta_k^2 a_k^2 \leq L^2$, and let $\phi_k$ be the cosine basis over $\Leb^2([-1,1])$, i.e.
+\begin{equation*}
+\phi_k(x) = 
+\begin{cases*}
+1/\sqrt{2},~~\textrm{for $k = 1$.} \,
+\cos\Bigl(\pi (k - 1) x\Bigr),~~\textrm{for $k \in \Nbb_1$}
+\end{cases*}
+\end{equation*}
+for any $x \in [-1,1]$. Put
+\begin{equation*}
+f_{\diamond} := \sqrt{2} \sum_{k = 1}^{\infty} \theta_k \phi_k.
+\end{equation*}
+Since $\sum_{k = 1}^{\infty} (\sqrt{2}\theta_k)^2 a_k^2 \leq 2L^2$, we know $f_{\diamond} \in H_{\mathrm{per}}^s([-1,1];2L)$. Moreover since $f_{\diamond}$ is an even function, we have that
+\begin{equation*}
+\sum_{\ell = 0}^{s}\int_{0}^{1} \bigl(f_{\diamond}^{(\ell)}(x)\bigr)^2 \,dx = \frac{1}{2} \sum_{\ell = 0}^{s}\int_{-1}^{1} \bigl(f_{\diamond}^{(\ell)}(x)\bigr)^2 \,dx \leq \frac{1}{\pi^{2s}}L^2
+\end{equation*}
+where the latter inequality follows from \textcolor{red}{(Tsybakov)}.
+As $f(x) = f_{\diamond}(x)$ for all $x \in [0,1]$, we have established that $f \in H^s([0,1],L')$ for $L' = L/\pi^s$. As for the boundary cases $x = 0$ and $x = 1$, putting
+\begin{equation*}
+f_N := \sum_{k = 1}^{N} \theta_k \varphi_k
+\end{equation*}
+we observe the following facts: first that $f_N^{(\ell)}(0) = f_N^{(\ell)}(1) = 0$ for all $N \in \Nbb$ and $\ell$ odd, and second that $f_N^{(\ell)}$ converges uniformly for all $0 \leq \ell < s$.
+Therefore we have $f^{(\ell)}(x) = \lim_{N \to \infty} f_N^{(\ell)}(x)$ for all $x \in [0,1]$ and $0 \leq \ell < s$, and in particular $f^{(\ell)}(0) = f^{(\ell)}(1) = 0$ for all $0 < \ell < s$ odd.
+
+Now for the $\supseteq$ direction. Let $f \in H^s([0,1];L)$ satisfy $f^{(\ell)}(0)$ for all $\ell = 1,\ldots,s - 1$ odd. Put
+\begin{equation*}
+f_{\diamond} = f(\abs{x})~~\textrm{for $x \in [-1,1]$}
+\end{equation*}
+and note the following facts: first that $f_{\diamond}^{(\ell)}(0)$ is well-defined for each $\ell = 0,\ldots, s - 1$, and second that $f_{\diamond}^{(s - 1)}$ is absolutely continuous on $[-1,1]$.
+As a result, $f_{\diamond}$ is $s$-times weakly differentiable (and $s - 1$ times classically differentiable), and by construction satisfies
+\begin{equation*}
+\norm{f_{\diamond}}_{H^s([-1,1])} = \sqrt{2} \norm{f}_{H^s([0,1])} \leq 2L
+\end{equation*} 
+Since clearly $f_{\diamond}^{(\ell)}(-1) = f_{\diamond}^{(\ell)}(1)$ for each $\ell = 0,\ldots, s - 1$, we have that $f_{\diamond} \in H_{\mathrm{per}}^s([-1,1];\sqrt{2}L)$.
+Since $f_{\diamond}$ is even, it may be expressed as $f_{\diamond} = \sum_{k = 1}^{\infty} \theta_k \phi_k$ where $\sum_{k = 1}^{\infty} \theta_k^2 a_k^{2} \leq 2L^2/\pi^{2s}$.
+Since $f(x) = f_{\diamond}(x)$ for $x \in [0,1]$, this implies
+\begin{equation*}
+f(x) = \sum_{k = 1}^{\infty} \theta_k \phi_k(x) = \sum_{k = 1}^{\infty} \frac{1}{\sqrt{2}}\theta_k \varphi_k(x) 
+\end{equation*}
+and the claim is proved.
 \end{document}

### 2020-03-12 notes/graph_testing_summary.tex
updated summary documents

@@
 
-We establish some notation that we will use throughout. Let $X = \{x_1,
-\ldots, x_n\}$ be a sample drawn i.i.d. from a distribution $P$ on $\Rd$,
-with density~$p$.  For a radius $r > 0$, we define $G_{n,r}=(V,E)$ to be the
-\emph{$r$-neighborhood graph} of $X$, an unweighted, undirected graph with
-vertices $V=X$, and an edge $(x_i,x_j) \in E$ if and only if $K_r(x_i,x_j) = \norm{x_i -x_j} \leq r$, where $\norm{\cdot}$ is the Euclidean norm. We denote by $A \in
-\Reals^{n \times n}$ the adjacency matrix, with entries $A_{uv} = 1$ if
-$(u,v) \in E$ and $0$ otherwise.  We also denote by $D$ the diagonal degree
-matrix, with entries $D_{uu} := \sum_{v \in V} A_{uv}$. The graph Laplacian is $L = D - A$, and we write its spectral decomposition as $L = V S V^T$.
+\section{Summary and Major TODOs}
 
-\section{Regression goodness-of-fit testing with random design.}
+Here is a list of results we've obtain so far. All rates are stated on the squared scale.
+\begin{itemize}
+	\item Assuming a random design regression model, the graph eigenvector projection test achieves the minimax optimal rate $n^{-4s/(4s + d)}$ over the Sobolev balls $H_0^s([0,1]^d;L)$ when $4s > d$.
+The restriction to compact support is unnecessary when $s = 1$.
+	\item Assuming a random design regression model, the graph eigenvector projection test achieves the minimax optimal rate $n^{-1/2}$ over the $\Leb^4$ ball $\Leb^4([0,1]^d;L)$. 
+	\item Assuming a fixed grid design regression model, the graph eigenvector projection test achieves the rate $n^{-4s/(4s + d)}$ over the Holder balls $C_0^s([0,1]^d;L)$ when $4s > d$.
+The restriction to compact support is unnecessary when $s = 1$
+	\item Assuming a fixed grid design regression model, the graph eigenvector projection test achieves the rate $n^{-2s/d}$ over the Holder balls $C_0^s([0,1]^d;L)$ for any $s$ and $d$.
+This is the minimax optimal rate when $4s \leq d$.
+	\item Assuming a fixed grid design regression model, the graph eigenvector projection test achieves the rate $n^{-4s/(4s + d)}$ over the Sobolev balls $\wt{H}^s([0,1]^d;L)$ when $2s > d$.
+	\item Simulation results for $d = 1$ showing that (a) when the density is uniform, the graph eigenvector projection test has similar testing error to the Fourier projection test, and the MMD using a Gaussian kernel, and (b) when the density is non-uniform and the regression function is spatially heterogeneous, the graph eigenvector projection test outperforms the latter two tests.
+\end{itemize}
 
-Let $P$ be a distribution with density $p$ supported on $\mathcal{X} \subseteq \Reals^d$. Suppose we observe random design points $X = \set{x_1,\ldots,x_n} \sim P$, and additionally responses
+Here is a list of results that are in progress.
+\begin{itemize}
+	\item All of the above statements apply equally to the Laplacian smoothing test.
+	\item Estimation results for the graph eigenvector projection and Laplacian smoothing estimators in $\Leb^2(P_n)$.
+\end{itemize}
+
+Here is a list of TODOs which I would say are high priority.
+\begin{itemize}
+	\item Defining ``out-of-sample'' estimators which agree with the graph eigenvector projection and Laplacian smoothing estimators at design points.
+Upper bounds on $\Leb^2(P)$ error of these estimators.
+	\item Upper bounds in testing and estimation when the data is assumed to lie on a manifold $\mathcal{M}$ with intrinsic dimension $m < d$.
+	\item Develop simulation results to include (a) the estimation problem, (b) dimensions $d > 1$, (c) more competitors.
+\end{itemize}
+Here is a list of TODOs which I would say are lower priority.
+\begin{itemize}
+	\item (Asymptotic) null distribution of our test statistics.
+	\item Investigate whether in the random design case, our assumptions on the smoothness of the density may be relaxed.
+	\item Show that for an appropriate modification of the Laplacian operator, our grid results can be stated without restricting the Sobolev space $H$ at the boundary.
+\end{itemize}
+
+\section{Preliminaries}
+
+We establish some notation that we will use throughout.
+A graph $G = (V,W)$ consists of a collection of vertices $V = \{v_1,\ldots,v_n\}$, and a weight matrix $W \in \Reals^{\abs{V} \times \abs{V}}$ with $(ij)$th entry $W_{ij}$ encoding affinity between $v_i$ and $v_j$.
+We will sometimes write $W_G$ to make it clear which graph a weight matrix is associated with.
+When the weight matrix $W_G$ consists only of $0$s and $1$s, we will use the equivalent representation $G = (V,E(G))$ when convenient; here $E(G) = V \times V \cap \{(v_i,v_j): W_{ij} = 1\}$.
+
+We will focus our attention on neighborhood graphs, and operators associated with them.
+For a kernel $K:\Reals \to \Reals$ and radius $r > 0$, we define $G_{n,K}=(V,W)$ to be the neighborhood graph with vertices $V = X$, and weights $W_{ij} = K(\norm{x_i - x_j}/r)$.
+Of particular interest to us will be the \emph{random geometric graph} $G_{n,r}$, the neighborhood graph constructed using the uniform kernel $K(z) := \1\{z \leq 1\}$.
+The weight matrix $W(G_{n,r})$ consists only of $0$s and $1$s, and we will thus use the equivalent representation $G_{n,r} = (X,E(G_{n,r}))$.
+
+The primary operator we are interested in (at this point) is the graph Laplacian. Let $D_G$ be the diagonal degree
+matrix of a graph $G$, with entries $D_{uu} := \sum_{v \in V} W_{uv}$.
+The graph Laplacian is then $L_G = D_G - W_G$.
+Let $\Lambda(G)$ consist of the eigenvalues $0 = \lambda_1(G) \leq \lambda_2(G) \leq \cdots \leq \lambda_n(G)$ of $L_G$, with $v_k(G) = (v_{k,1}(G),\ldots,v_{k,n}(G)) \in \Reals^n$ denoting the eigenvector corresponding to the $k$th eigenvalue $\lambda_k(G)$.
+
+\section{Regression goodness-of-fit testing.}
+
+Suppose we observe samples $(y_i,x_i)$ for $i = 1,\ldots,n$, where conditional on $X$ the responses $Y = \{y_1,\ldots,y_n\}$ are assumed to follow the model
 \begin{equation}
@@
 \paragraph{Eigenvector projection test statistic.}
-Let $VSV^T$ be the spectral decomposition of the Laplacian matrix $L$ of the neighborhood graph $G_{n,r}$.
-To test whether $f = f_0$, we propose the following \emph{eigenvector projection} test statistic:
+The graph Laplacian eigenvector projection test is a truncated-series test.
+Suppose we are given a graph $G = (X,W)$ defined on the sample points $X$.
+Letting $\kappa$ be some integer between $1$ and $n$, our test projects $Y$ onto the eigenvectors $v_1(G),\ldots,v_{\kappa}(G)$, and then takes the empirical norm of this projection to be the test statistic.
+Formally, let
+\begin{equation*}
+\Pi_{\kappa,G}(f) := \frac{1}{n}\sum_{k = 1}^{\kappa} \Biggl(\sum_{i = 1}^{n} f(x_i) v_{k,i}(G)\Biggr) v_{k}(G),~~ T_{\mathrm{spec}}(G) := \norm{\Pi_{\kappa,G}(f)}_n^2
+\end{equation*}
+where $\norm{\theta}_n^2 := \frac{1}{n} \sum_{i = 1}^{n} \theta_i^2$ for $\theta \in \Reals^n$.
+Then the test $\phi_{\spec}(G) := \1\{T_{\spec}(G) \geq \tau\}$ rejects the null hypothesis when the test statistic $T_{\mathrm{spec}}(G)$ is greater than a pre-specified cutoff $\tau$.
+
+\subsection{Random design.}
+In the random design case, we assume that each design point $x_i$ is sampled independently from some distribution $P$. Formally, our model for the samples is the following:
 \begin{equation}
-\label{eqn:graph_spectral_projections}
-T_{\mathrm{spec}} := \frac{1}{n} \sum_{k = 0}^{\kappa} \left(\sum_{i = 1}^{n} v_{k,i} y_i\right)^2
+\label{eqn:regression_random_design_known_variance}
+x_1,\ldots,x_n \overset{\textrm{i.i.d}}{\sim} P,~~ \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)~\textrm{and}~ \varepsilon \perp X,~~ y_i = f(x_i) + \varepsilon_i.
 \end{equation}
 
-\subsection{Current Results.}
+In Theorem~\ref{thm:sobolev_testing_rate_order1} we show that under some typical regularity conditions on $P$, the neighborhood graph Laplacian eigenvector projection test $\phi_{\textrm{spec}}(G_{n,r})$ is a minimax optimal test over the Sobolev ball $H^1(\mathcal{X};R)$ for $d = 1,2,3$ and $\Xset = [0,1]^d$.
+Let $p_d$ equal $3/4$ when $d = 2$ and equal $1/d$ for $d \geq 3$.
 
-In Theorem~\ref{thm:sobolev_testing_rate} we show that under some typical regularity conditions on $P$, the test $\phi_{\textrm{spec}} := \1\{T_{\mathrm{spec}} \geq \tau\}$ is, up to log factors, a minimax optimal test over the Sobolev ball $W^{1,2}(\mathcal{X};R)$.
-To conveniently state our results we introduce the notation
+\begin{theorem}
+	\label{thm:sobolev_testing_rate_order1}
+	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}.
+Let $R > 0$ and $b \geq 1$ be fixed constants, and $d = 1,2$ or $3$.
+Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p(x)$ bounded away from zero and infinity,
 	\begin{equation*}
-h(a,d) = a(d+2) + (1 + 2/d),~~\textrm{for $a > 0$}
+	0 < p_{\min} < p(x) < p_{\max} < \infty,~~ \textrm{for all $x \in \mathcal{X}$.}
 	\end{equation*}
+	and the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choices
+	\begin{equation*}
+	c \frac{(\log n)^{p_d}}{n^{1/d}} \leq r(n) \leq n^{-4/((4 + d)(2+d))}, ~\kappa = (nR)^{2d/(4 + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	\end{equation*}
+	for $c$ a constant that depends only on $\Xset, p_{\min}$ and $p_{\max}$.
+Then the following statements holds for every $n$ sufficiently large: there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
+	\begin{equation}
+	\label{eqn:sobolev_testing_rate_order1}
+	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4 + d)}\} \cdot n^{-4/(4 + d)}
+	\end{equation}
+	the worst-case risk is upper bounded
+	\begin{equation}
+	\label{eqn:sobolev_testing_rate_order1_1}
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}(G_{n,r}); H^1(\mathcal{X};R)) \leq \frac{c_2}{b}.
+	\end{equation}
+\end{theorem}
+
+The testing problem in the random design case with known variance fundamentally changes in the low-smoothness regime, where the number of bounded derivatives is small compared to the dimension $d$.
+In the case where $f \in H^1(\Xset;R)$ is assumed to have one derivative which is bounded in $\Leb^2(\Xset)$ norm, this transition occurs at $d = 4$.
+We shall discuss testing in the low-smoothness regime shortly.
+First, let's consider the situation when we are willing assume $f$ possesses additional derivatives bounded in $\Leb^2$ norm.
+
+Theorem~\ref{thm:sobolev_testing_rate_order1} establishes that a truncated series test using graph eigenvectors performs just as well, in a minimax sense, as a truncated series test using a classic Fourier series; for analysis of the latter test as well as derivation of the minimax rate of nonparametric goodness-of-fit testing over Sobolev classes see \textcolor{red}{(Ingster)}.
+Intuitively, we can understand this by thinking of the graph eigenvectors as estimates of eigenfunctions of a density-weighted continuum Laplacian operator.
+When the density is uniform, and the domain $\Xset$ is the unit cube, under appropriate boundary conditions these eigenfunctions are themselves elements of the Fourier basis.
+When the density is merely close to uniform, the eigenfunctions no longer belong to the Fourier basis; nevertheless, they share enough properties with the latter that the resulting minimax rate is unchanged.
+Theorem~\ref{thm:sobolev_testing_rate_order1} shows that the additional error due to using \textit{estimates} of these eigenfunctions does not fundamentally change the minimax rate.
+
+When we are additionally willing to assume $H^s(\Xset)$, we know (again from \textcolor{red}{Ingster}) that the minimax testing rate is $n^{-4s/(4s + d)}$.
+Now, however, our estimates of the eigenvectors no longer track the derivatives finely enough at the boundaries of $\Xset$ to establish the same minimax rate over all $H^s(\mathcal{X})$.
+However, if we assume $f$ and its derivatives are compactly supported on $\Xset$, we recover the usual rate.
+Theorem~\ref{thm:sobolev_testing_rate} presents our formal result, that $\phi_{\textrm{spec}}(G_{n,r})$ is a minimax optimal test over the Sobolev balls $H_0^s(\mathcal{X};R)$ whenever $4s > d$ and $\Xset = [0,1]^d$.
 
@@
 	\label{thm:sobolev_testing_rate}
-	Let $b \geq 1$ and $a > 0$ be fixed constants, and let $d < 4$.
-Suppose that $\Pbb$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p$ bounded above and below by constants, i.e
+	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}.
+Let $R > 0$, $b,s,d \geq 1$ be fixed constants, with $s$ and $d$ integers.
+Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p \in C^{s-1}(\Xset;p_{\max})$ for some $p_{\max} < \infty$, and further that $p(x)$ is bounded away from zero, i.e.
+there exists $p_{\min} > 0$ such that
 	\begin{equation*}
-	0 < p_{\min} < p(x) < p_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
+	p_{\min} < p(x),~~ \textrm{for all $x \in \mathcal{X}$.}
 	\end{equation*}
-	Then the following statement holds: if the test $\phi_{\spec}$ is performed with parameter choices 
+	Then the following statement holds for all $n$ sufficiently large: if the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choice
 	\begin{equation*}
-	r = \log^a n \cdot \left(\frac{\log n}{n}\right)^{1/d}, ~\kappa = n^{2d/(4 + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	n^{-1/(2(s-1) + d)} \leq r(n) \leq n^{-4/((4s + d)(2+d))}, ~\kappa = (Rn)^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
 	\end{equation*}
-	then there exists a constant $c$ which may depend on $d,R,p_{\max},b$ and $a$ but is independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
+	then there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
 	\begin{equation}
 	\label{eqn:sobolev_testing_rate}
-	\epsilon^2 \geq c \cdot b \cdot n^{-4/(4 + d)} (\log n)^{h(a,d)}
+	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot \max\{R^2,R^{2d/(4s + d)}\} \cdot n^{-{4s}/(4s + d)}
 	\end{equation}
@@
 	\label{eqn:sobolev_testing_rate_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}.
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; H_0^{s}(\mathcal{X};R)) \leq \frac{c_2}{b}.
 	\end{equation}
@@
 
-When $d \geq 4$ the compact embedding
+The restriction that $p$ be supported on the unit cube is mostly for convenience.
+If instead $p$ is supported on any compact set with Lipschitz boundary, the Theorem statement should hold (up to constants), with only a few modifications to our proofs.
+However, we do not work through the details.
+
+The restriction $4s > d$ is a fundamental consequence of the tightness of the Sobolev embedding theorem. When $4s \geq d$ the compact embedding
 \begin{equation*}
-W_d^{1,2}(\mathcal{X}) \subseteq \mathcal{L}_d^4(\mathcal{X}) 
+H^{s}(\mathcal{X}) \subseteq \Leb^4(\mathcal{X}) 
 \end{equation*}
-does not hold (as it does when $d < 4$). However, if we directly assume $f \in \mathcal{L}_d^4(\mathcal{X};R)$ (regardless of $d$) we obtain the following result.
+does not hold (as it does when $d < 4s$). However, if we directly assume $f \in \mathcal{L}_d^4(\mathcal{X};R)$ (regardless of $d$) we obtain the following result.
 \begin{proposition}
 	\label{prop:L4_testing_rate}
-	If $f \in \mathcal{L}_d^4(\mathcal{X};R)$, there exists a constant $c$ such that if
+	If $f \in \Leb^4(\mathcal{X};R)$, there exists a constant $c$ such that if
 	\begin{equation}
 	\label{eqn:L4_testing_rate}
-	\epsilon^2 > b \cdot n^{-1/2}
+	\epsilon^2 > c^2 b \cdot n^{-1/2}
 	\end{equation}
@@
 	\begin{equation*}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; W^{1,2}(\mathcal{X};R)) \lesssim \frac{1}{b}. 
+	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; H^1(\mathcal{X};R)) \lesssim \frac{1}{b}. 
+	\end{equation*}
+\end{proposition}
+Note that when $4s > d$ the critical radius~\eqref{eqn:sobolev_testing_rate} is smaller than \eqref{eqn:L4_testing_rate}.
+
+\subsection{Fixed Design}
+Frequently in nonparametric regression problems, it is assumed that the design points are fixed, and, in the simplest case, uniformly spaced on the unit cube.
+For convenience in this setting we let $n = N^d$ for some integer $N$.
+The fixed grid design $\wb{X}$ consists of $n$ total evenly spaced grid points on $[0,1]^d$,
+\begin{equation*}
+\wb{X} := \Bigl\{\Bigl(\frac{2i_1 - 1}{2N},\ldots,\frac{2i_d - 1}{2N}\Bigr): i \in [N]^d\Bigr\}
+\end{equation*}
+Suppose we observe $n$ samples according to regression model
+\begin{equation}
+\label{eqn:grid_regression_model}
+y_i = f(\wb{x}_i) + \varepsilon_i, ~\varepsilon_i \overset{\mathrm{i.i.d}}{\sim} \mathcal{N}(0,1) ~~\textrm{for each $i \in [N]^d$,}
+\end{equation}
+Clearly model~\eqref{eqn:grid_regression_model} bears some similarity with model~\eqref{eqn:regression_random_design_known_variance}, the main difference being that the design $\wb{X}$ is now fixed instead of random.
+
+We study the behavior of $\phi_{\spec}(\wb{G})$, where $\wb{G}$ is the lattice
+\begin{equation*}
+\wb{G} := \bigl(\wb{X},\wb{E}\bigr),~~ \wb{E} := \set{(\wb{x}_i,\wb{x}_j): i,j \in [N]^d, \norm{i - j}_1 = 1},
+\end{equation*}
+a very natural graph to build given our assumed grid design. Note that for a given $d$, we also have the recursive relation
+\begin{equation*}
+\wb{G} := \wb{G}_1 \otimes \ldots \otimes \wb{G}_1
+\end{equation*}
+where the tensor product is taken $d$ total times, and $\wb{G}_1$ is the $1$d lattice (i.e.
+the path) graph on $N$ vertices.
+For this reason it will be convenient to change notation slightly, and index by $d$-tuples of numbers rather than by numbers.
+In particular, we index the eigenvalues and eigenvectors of $L_{\wb{G}}$ as
+\begin{equation*}
+\lambda_k(\wb{G}) = \prod_{j = 1}^{d} \lambda_{k_j}(\wb{G}_1), v_{k,i}(\wb{G}) = \prod_{j = 1}^{d} v_{k_j,i_j}(\wb{G}_1)~~\textrm{for $k,i \in [N]^d$.}
+\end{equation*}
+Then, for a given $\kappa \in [n]$, letting $K = \kappa^{1/d}$, we see that our graph spectral test statistic may be written equivalently as
+\begin{equation*}
+T_{\spec}(\wb{G}) = \frac{1}{n} \sum_{k \in [K]^d} \Biggl(\sum_{i \in [N]^d} y_i v_{k,i}(G)\Biggr)^2
+\end{equation*}
+
+We see the test $\phi_{\spec}(\wb{G})$ achieves the same rate over the Holder ball $C^{1}(\Xset;L)$ when $d \leq 4$, as the test $\phi_{\spec}(G_{n,r})$ does over the Holder ball $C^1(\Xset;L)$.
+\begin{theorem}
+	\label{thm:holder_testing_rate_grid}
+	Let $L > 0$ be a fixed constant, and let $d \leq 4$.
+Suppose we observe data according to the model~\eqref{eqn:grid_regression_model}.
+Then when the test $\phi_{\spec}(\wb{G})$ is performed with the parameter choices
+	\begin{equation*}
+	\kappa = (Ln)^{2d/(4 + d)},~~ \tau(b) = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	\end{equation*}
+	for some $b \geq 1$, the following statement is true: there exists a constant $c$ such that for 
+	\begin{equation}
+	\label{eqn:holder_testing_rate_grid_1}
+	\epsilon^2 \geq c \cdot \max\{L^{d/(4+d)}, L^{d/(4 + d) + 1}\} \cdot b n^{-4/(4 + d)}
+	\end{equation}
+	the worst-case risk of $\phi_{\spec}(\wb{G})$ on $C^{1}(\Xset;L)$ is upper bounded
+	\begin{equation}
+	\label{eqn:holder_testing_rate_grid_2}
+	\mathcal{R}_{\epsilon}\Bigl(\phi_{\spec}(\wb{G});C^{1}(\Xset;L)\Bigr) \leq \frac{4}{b^2} + \frac{8}{b\sqrt{2\kappa}}.
+	\end{equation}
+\end{theorem}
+
+As was the case with the neighborhood graph in the random design setting, the grid Sobolev norm fails to track the higher order derivatives of $f$ near the boundary of $\Xset$.
+We therefore will again insist that $f$ be compactly supported within $\Xset$.
+Under this condition, we find that the test $\phi_{\spec}(\wb{G})$ achieves the same rate over the Holder ball $C_c^{s}(\Xset;L)$ when $4s \geq d$ as did the test $\phi_{\spec}(G_{n,r})$ over the Holder ball $C_c^s(\Xset;L)$.
+
+\begin{theorem}
+	\label{thm:holder_testing_rate_grid_higher_order}
+	Let $L > 0$ be a fixed constant, $s \geq 2$ be a fixed integer, and let $d \leq 4s$.
+Suppose we observe data according to the model~\eqref{eqn:grid_regression_model}.
+Then when the test $\phi_{\spec}(\wb{G})$ is performed with the parameter choices
+	\begin{equation*}
+	\kappa = (Ln)^{2d/(4s + d)},~~ \tau(b) =  \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	\end{equation*} 
+	for some $b \geq 1$, the following statement is true: there exists a constant $c$ such that for any
+	\begin{equation*}
+	\epsilon^2 \geq c L^2 b n^{-4s/(4s + d)},
+	\end{equation*}
+	the worst-case risk of $\phi_{\spec}(\wb{G})$ is upper bounded
+	\begin{equation*}
+	\mathcal{R}_{\epsilon}\Bigl(\phi_{\spec}(\wb{G});C_c^{s}(\Xset;L)\Bigr) \leq \frac{4}{b^2} + \frac{8}{b\sqrt{2\kappa}}.
+	\end{equation*}
+\end{theorem}
+
+Roughly speaking, we have shown that in the high-smoothness regime $4s > d$, testing over Holder classes in the grid design and random design cases is an equally difficult problem from the minimax perspective.
+The same is not true for the corresponding Sobolev spaces.
+When $2s > d$, the Sobolev spaces $H^s(\Xset;L)$ embed compactly into (lower-order) Holder classes, implying that every equivalence class $u \in H^s(\Xset;L)$ contains a favored representer which is bounded and Holder continuous, with bounded Holder norm.
+When $2s < d$, the situation is less nice; the spaces $H^s(\Xset;L)$ now contain (equivalence classes of) bump functions of arbitrarily large height and small width.
+By placing such bump functions at design points, it is possible to make the sampling model~\eqref{eqn:grid_regression_model} contain arbitrarily little information about the overall behavior of the function $f$ over $\Xset$.
+It is not hard to construct a function $f \in H^s(\Xset;L)$ for which, say, $\norm{f}_{\Leb^2(\Xset)} = 1$ yet $f(\wb{x}) = 0$ for all $\wb{x} \in \wb{X}$, implying that the critical radius $\epsilon^{\star}(H^s(\Xset;L)) = \Omega(1)$.
+We note that this is not an issue in the random design setting, because the design points are (randomly) chosen only after the function $f$ is fixed.
+
+\textcolor{red}{(TODO): Make this discussion more rigorous. Perhaps appeal to the concept of precise representatives defined in (Evans and Gariepy).}
+
+The above discussion leaves hope that when $2s > d$, consistent testing might still be possible.
+We can in fact go beyond this and show that the graph spectral test $\phi_{\spec}(\wb{G})$ achieves the same rate over the Sobolev space $\wt{H}^s(\Xset;L)$ as the test $\phi_{\spec}(G_{n,r})$ did over $H_0^s(\Xset;L)$.
+
+\begin{theorem}
+	\label{thm:sobolev_testing_rate_grid}
+	Let $L > 0$ be a fixed constant, $s \geq 1$ be a fixed integer, and let $2s > d$.
+Suppose we observe data according to the model~\eqref{eqn:grid_regression_model}.
+Then when the test $\phi_{\spec}(\wb{G})$ is performed with the parameter choices
+	\begin{equation*}
+	\kappa = n^{2d/(4s + d)},~~ \tau(b) =  \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	\end{equation*} 
+	for some $b \geq 1$, the following statement is true: there exists a constant $c$ such that for any
+	\begin{equation*}
+	\epsilon^2 \geq c L^2 b n^{-4s/(4s + d)},
+	\end{equation*}
+	the worst-case risk of $\phi_{\spec}(\wb{G})$ is upper bounded
+	\begin{equation}
+	\label{eqn:sobolev_testing_rate_grid_1}
+	\mathcal{R}_{\epsilon}\Bigl(\phi_{\spec}(\wb{G});\wt{H}^s(\Xset;L)\Bigr) \leq \frac{4}{b^2} + \frac{8}{b\sqrt{2\kappa}}
+	\end{equation}
+	for all $n$ sufficiently large.
+\end{theorem}
+
+\subsubsection{Low-smoothness regime with fixed design}
+In the low-smoothness regime $d > 4s$, the coupling between empirical norm over $\wb{X}$ and the continuum $\Leb^2$ norm is sufficiently weak that the discretization error can become the dominant source of error, as we show in Lemma~\ref{lem:holder_testing_rate_grid_low_smoothness_lb}.
+
+\begin{lemma}
+	\label{lem:holder_testing_rate_grid_low_smoothness_lb}
+	For any $L > 0$ and any integers $s$ and $d$, there exists a function $g \in C^s(\Xset;L)$ such that
+	\begin{equation*}
+	\norm{g}_{\Leb^2}^2 \geq \frac{L^2}{2^d \pi^{2s} (d + 1)^{2s}}n^{-2s/d} 
+	\end{equation*}
+	but $g(\wb{x}) = 0$ for all $\wb{x} \in \wb{X}$. 
+\end{lemma}
+Lemma~\ref{lem:holder_testing_rate_grid_low_smoothness_lb} implies the critical radius over the Holder classes $C^s(\Xset;L)$ must be at least on the order of $\epsilon^{\star}(C^s(\Xset;L)) \gtrsim n^{-s/d}$.
+When $4s \geq d$, this term is neglible relative to $n^{-2s/(4s + d)}$, and we arrive at the ``usual'' rates of minimax testing we see in Theorems~\ref{thm:holder_testing_rate_grid} and~\ref{thm:holder_testing_rate_grid_higher_order}.
+On the other hand when $4s < d$, this becomes the dominant term, and we can no longer achieve the typical rate $\epsilon^2 \asymp n^{-4s/(4s + d)}$.
+
+In this setting, when we enforce the wider radius $\epsilon^2 \asymp n^{-2s/d}$, Proposition~\ref{prop:holder_testing_rate_grid_low_smoothness_ub} shows that a very simply test has small worst-case risk $\mathcal{R}_{\epsilon}$.
+
+\begin{proposition}
+	\label{prop:holder_testing_rate_grid_low_smoothness_ub}
+	Let $L > 0$ be a fixed constant, $s \geq 1$ a fixed integer, and let $d \geq 4s$.
+Suppose we observe data according to the grid design regression model~\eqref{eqn:grid_regression_model}.
+Then when the test $\phi_{\spec}(\wb{G})$ is performed with the parameter choices
+	\begin{equation*}
+	\kappa = n,~~ \tau(b) = 1 + b\sqrt{\frac{2}{n}}
+	\end{equation*}
+	for some $b \geq 1$, the following statement is true: there exists a constant $c$ such that for any 
+	\begin{equation*}
+	\epsilon^2 \geq c b L^2 n^{-2s/d}
+	\end{equation*}
+	the worst case risk of $\phi_{\spec}(\wb{G})$ is upper bounded
+	\begin{equation*}
+	\mathcal{R}_{\epsilon}(\phi_{\spec}(\wb{G}); C^s(\Xset;L)) \leq \frac{4}{b^2} + \frac{8}{b\sqrt{2n}}
 	\end{equation*}
@@
 
-Note that when $d < 4$ the rate~\eqref{eqn:sobolev_testing_rate} is sharper than \eqref{eqn:L4_testing_rate}.
+\begin{itemize}
+	\item The test statistic $T_{\spec}(\wb{G})$ is simply the empirical norm of $Y$. 
+	\item Together, Lemma~\ref{lem:holder_testing_rate_grid_low_smoothness_lb} and Proposition~\eqref{prop:holder_testing_rate_grid_low_smoothness_ub} characterize the minimax rate of the grid design regression testing problem when $4s > d$.
+	\item This minimax rate does not match the upper bound established in Proposition~\ref{prop:L4_testing_rate}---the fixed design problem is harder. 
+	\item Note that the usual minimax rate of estimation error over Holder classes is always $n^{-2s/(2s + d)}$, regardless of the relation between $s$ and $d$, and regardless of whether loss is measured in $\Leb^2$-norm or empirical norm.
+The explanation for this is that the rate $n^{-2s/(2s + d)}$ is always larger than $n^{-2s/d}$, and so the discretization error is never the dominant source of error in estimation.
+This reveals an interesting distinction between the testing and estimation problems, as testing is easy enough in a statistical sense that error incurred by the discrete nature of the problem may be the bottleneck.
+\end{itemize}
 
-The minimax testing rate over the higher-order smoothness classes $W_d^{s,2}(\Xset;L)$ and $C_d^s(\Xset;L)$ is known to be $n^{-4s/(4s + d)}$ when $4s < d$.
-Hence the testing rate~\eqref{eqn:sobolev_testing_rate} matches the minimax critical radius when $s = 1$ up to a factor of $\log^{h(a,d)}n$.
-The eigenvector projection test is also minimax optimal over the balls in higher order Holder spaces $C_d^s(\Xset;L)$.
+\subsection{Analysis}
 
-\begin{theorem}
-	\label{thm:higher_order_holder_testing_rate}
-	Let $b \geq 1$ be a fixed constant, and let $d$ and $s$ be positive integers such that $d < 4s$.
-Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p \in C^{s-1}(\mathcal{X};R)$ bounded above and below by constants, i.e
-	\begin{equation*}
-	0 < p_{\min} < p(x) < p_{\max} < \infty, \quad \textrm{for all $x \in \mathcal{X}$.}
-	\end{equation*}
-	Then the following statement holds: if the test $\phi_{\spec}$ is performed with parameter choices 
-	\begin{equation*}
-	\textrm{$K_r$ the uniform kernel with}~ n^{-1/(2(s - 1) + d)} \leq r(n) \leq n^{-4/((4s + d)(2+d))}, ~\kappa = n^{2d/(4s + d)}, ~\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
-	\end{equation*}
-	then there exists constants $c_1,c_2$ which may depend on $d,R,$ and $s$ but are independent of the sample size $n$ such that for every $\epsilon \geq 0$ satisfying
+To prove Theorem~\ref{thm:sobolev_testing_rate}, we show that there exists a high-probability set $E \subseteq \Xset^n$ such that conditional on $X \in E$, the test $\phi_{\spec}(G_{n,r})$ has small risk.
+Since $G_{n,r}$ is a function only of $X$ and not of $Y$, this amounts to reasoning about the behavior of the test $\phi_{\spec}$ over a fixed graph $G = (X,E)$, where we observe
 \begin{equation}
-	\label{eqn:higher_order_holder_testing_rate}
-	\epsilon^2 \geq c_1 \cdot b^2 \cdot n^{-4s/(4s + d)}
+\label{eqn:fixed_graph_regression_model}
+y_i = \beta_i + \varepsilon_i,~~\varepsilon_i \sim \mathcal{N}(0,1)
 \end{equation}
-	the worst-case risk is upper bounded
+for some fixed $\beta \in \Reals^n$.
+In Lemma~\ref{lem:fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}(G)$.
+Our bound on the Type II error will be stated as a function of $\beta^T L^s \beta$--a measure of the smoothness the signal $\beta$ displays over the graph $G$--as well as the $\kappa$th eigenvalue $\lambda_{\kappa}$.
+
+\begin{lemma}
+	\label{lem:fixed_graph_testing}
+	Let $1 \leq \kappa \leq n$ be an integer.
+Suppose we observe data according to the model~\eqref{eqn:fixed_graph_regression_model}, and perform the test $\phi_{\spec}(G)$ with $\tau(b) = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}$.
+	\begin{enumerate}
+		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}(G)$ is upper bounded
 		\begin{equation}
-	\label{eqn:higher_order_holder_testing_rate_1}
-	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}; C_d^{s}(\mathcal{X};R)) \leq \frac{c_2}{b}.
+		\label{eqn:graph_spectral_type_I_error}
+		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{2}{b^2}.
 		\end{equation}
-\end{theorem}
+		\item \textbf{Type II error:} For any $b \geq 1$ and $\beta$ such that
+		\begin{equation}
+		\label{eqn:fixed_graph_testing_critical_radius_1}
+		\norm{\Pi_{\kappa,G}(\beta)}_n^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}}
+		\end{equation}
+		the Type II error of $\phi_{\spec}(G)$ is upper bounded,
+		\begin{equation}
+		\label{eqn:graph_spectral_type_II_error}
+		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{2}{b^2} + \frac{8}{b\sqrt{2\kappa}}.
+		\end{equation}
+		In particular if
+		\begin{equation}
+		\label{eqn:fixed_graph_testing_critical_radius}
+		\frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{\beta^T L^s \beta}{n\lambda_{\kappa}^s}
+		\end{equation}
+		then~\eqref{eqn:fixed_graph_testing_critical_radius_1} and thus~\eqref{eqn:graph_spectral_type_II_error} follow.
+	\end{enumerate}
+\end{lemma}
 
@@
 \begin{equation*}
-\norm{f}_{C_d^{s}(\mathcal{X})} := \sum_{\abs{\alpha} < s} \norm{D^{\alpha}f}_{\infty} + \sum_{\abs{\alpha} = \ell} \sup_{x,y \in \mathcal{X}} \frac{\abs{D^{\alpha}f(y) - D^{\alpha}f(x)}}{\norm{x - y}_2^{s - \ell}}
+\norm{f}_{C^{s}(\mathcal{X})} := \sum_{\abs{\alpha} < s} \norm{D^{\alpha}f}_{\infty} + \sum_{\abs{\alpha} = \ell} \sup_{x,y \in \mathcal{X}} \frac{\abs{D^{\alpha}f(y) - D^{\alpha}f(x)}}{\norm{x - y}_2^{s - \ell}}
 \end{equation*}
-and the $s$th Holder space $C_d^{s}(\mathcal{X})$ consists of all functions which are $\ell$ times continuously differentiable with finite Holder norm $\norm{\cdot}_{C_d^{s}(\mathcal{X})}$.
-Denote the Holder ball by $C_d^{s}(\mathcal{X},R) = \set{f \in C_d^{s}(\mathcal{X}): \norm{f}_{C_d^{s}(\mathcal{X})} \leq R}$.
+and the $s$th Holder space $C^{s}(\mathcal{X})$ consists of all functions which are $\ell$ times continuously differentiable with finite Holder norm $\norm{\cdot}_{C^{s}(\mathcal{X})}$.
+Denote the Holder ball by $C_d^{s}(\mathcal{X},R) = \set{f \in C^{s}(\mathcal{X}): \norm{f}_{C_d^{s}(\mathcal{X})} \leq R}$.
+Let $C_c^{s}(\Xset)$ consist of those functions in $C^s(\Xset)$ which are compactly supported within $\Xset$; formally $f \in C_c^{s}(\Xset)$ if and only if $f \in C^s(\Xset)$ and there exists an open set $V \subset \ol{V} \subset \Xset$ such that $\mathrm{supp}(f) \subset V$.
 
-\paragraph{Sobolev spaces.}
+\subsubsection{Sobolev spaces}
 
-For a given $s > 0$, the Sobolev space $W_d^{s,2}(\mathcal{X})$ consists of all functions $f \in \mathcal{L}^2(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^2(\mathcal{X})$.
-The Sobolev $\{s,2\}$ norm is then
+For a given $s$ and $p > 0$, the Sobolev space $W^{s,p}(\mathcal{X})$ consists of all functions $f \in \Leb^2(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^2(\mathcal{X})$.
+The Sobolev $\{s,p\}$ norm is then
 \begin{equation*}
-\norm{f}_{W_d^{s,2}(\mathcal{X})}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^2 \,dx
+\norm{f}_{W^{s,p}(\mathcal{X})}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^p \,dx
 \end{equation*}
-and for a given $L > 0$, the corresponding ball is $W_d^{s,2}(\Xset; L) = \set{f: \norm{f}_{W^{s,2}(\Xset)} \leq L}$.
+and for a given $R > 0$, the corresponding ball is $W^{s,p}(\Xset; R) = \set{f: \norm{f}_{H^s(\Xset)} \leq R}$.
+In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space; we adopt the usual convention of writing $H^s(\Xset) = H^s(\Xset)$.
+
+We write $W_0^{s,p}(\Xset)$ for the completion of $C_c^{\infty}(\Xset)$, and let $H_0^s(\Xset) = W_0^{s,2}(\Xset)$.
+
+\paragraph{Defining Sobolev spaces using Fourier series.}
+
+Let the cosine basis on $\Leb^2([0,1])$ be defined as 
+\begin{equation*}
+\varphi_k(x) = 
+\begin{cases*}
+1,~~\textrm{for $k = 1$.} \,
+\sqrt{2}\cos\Bigl(\pi (k - 1) x\Bigr),~~\textrm{for $k \in \Nbb_1$}
+\end{cases*}
+\end{equation*}
+for any $x \in [0,1]$. Let the tensor product cosine basis on $\Leb^2([0,1]^d)$ be defined as
+\begin{equation*}
+\varphi_k(x) = \prod_{j = 1}^{d} \varphi_{k_j}(x_j), ~~\textrm{for $k \in \Nbb^d$},
+\end{equation*}
+for any $x \in [0,1]^d$. Observe that $\varphi_k(\wb{x}_i) = \sqrt{n} v_{k,i}(\wb{G}_d)$ for all $k,i \in [N]^d$.
+
+We let the space 
+\begin{equation*}
+\wt{H}^{s}(\Xset;L) = \Bigl\{ \sum_{k \in \mathbb{N}^d} \theta_k \varphi_k: \sum_{k \in \mathbb{N}^d} \theta_k^2 a_k^{2} \leq L \Bigr\}
+\end{equation*}
+where in the one-dimensional case $a_{k} = (k - 1)^s$ and for general $d$, $a_k = \sqrt{\sum_{j = 1}^{d} a_{k_j}^2}$.
+It is not hard to show that, when $s$ is an integer, $\wt{H}^s([0,1]^d)$ consists of those functions in $H^s([0,1]^d)$ which satisfy some Neumann-type boundary conditions.
+For example when $d = 1$,
+\begin{equation*}
+\wt{H}^s([0,1];L) = \Bigl\{f \in H^{s}([0,1];L \pi^{s}): f^{(\ell)}(0) = f^{(\ell)}(1) = 0,~~\textrm{for all odd integers $0 < \ell < s$} \Bigr \} \supseteq H_0^s([0,1];L'),
+\end{equation*}
+a fact which we prove in Section~\ref{subsec:sobolev_class_equivalence}.
 
@@
 
+\subsection{Kernels}
+We say a kernel function $K(\cdot)$ is a $2$nd order kernel if $K$ is compactly supported on $B(0,1)$, uniformly upper bound $\abs{K(x)} \leq K_{\max} < \infty$ for all $x \in B(0,1)$, and
+\begin{equation*}
+\int K(x) \,dx = \nu_d,~~ \int x K(x) \,dx = 0.
+\end{equation*}
+Note that the uniform kernel, defined as $K(x) = \1\{\norm{x} \leq 1\}$, is a $2$nd order kernel.
+
+\subsection{Notation}
+
+We will treat $f$ interchangeably as a function $f:\Rd \to \Reals$, and as a vector $f = (f(x_1),\ldots,f(x_n))$.
+More generally, for $U \subset V$ and $f:V \to \Reals$, the action of $f$ on $U$ is naturally defined via the restriction operator $R_U$; we will forego additional notation and simply treat $f$ as a function on $U$, defined as $f(x) = R_Uf(x)$ for $x \in U$.
+It should always be clear from context how we are using $f$.
+
+For sequences $(a_n), (b_n)$, we say $a_n = O(b_n)$ if there exists a constant $c$ such that $a_n \leq cb_n$ for all $c$. We use the notation $a \wedge b$ for the minimum of $a$ and $b$.
+
+We use the notation $\Lambda(H)$ to denote the spectrum of a matrix $H$, and $\lambda_k(H)$ to denote the $k$th smallest eigenvalue of $H$.
+
+We sometimes -- and currently not in a consistent manner -- denote the lattice Laplacian $L_{\wb{G}} =: \wb{L}$.
+
+We write $\mathbb{N}_k$ for the natural numbers excluding $0,\ldots,k - 1$. We write $\mathbb{N}_0$ for the natural numbers including $0$, and $\mathbb{N} = \mathbb{N}_1$.
+
 \end{document}

### 2020-08-04 notes/graph_regression2.tex
First commit at second stab at paper on graph regression. Added section on estimation with Laplacian eigenmaps.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{tikz}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}
\newcommand{\Lap}{{\bf L}}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Minimax-optimal learning with graph Laplacians}
\author{Alden Green}
\date{\today}
\maketitle

\section{Introduction}

\textcolor{red}{TODO:} Paragraph motivating graph Laplacian methods.

In the random design nonparametric regression setup, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ on a domain $\Xset \subset \Rd$ and 
\begin{equation}
\label{eqn:random_design_regression}
Y_i = f_0(X_i) + \varepsilon_i
\end{equation}
with independent Gaussian noise $\varepsilon_i \sim N(0,1)$. The goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether or not $f_0 = 0$, i.e whether or not there is a signal present.

In graph based learning, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$, which should approximately capture the geometry of $P$ and $\mc{X}$, and then constructing an estimate $\wh{f}$ which is smooth with respect to $G_{n,r}$. For a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$. The weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ is an $n \times n$ matrix with entries
\begin{equation*}
\label{eqn:neighborhood_graph}
{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr),
\end{equation*}
and the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$.  Our star operator, the neighborhood graph Laplacian, can then be written as
\begin{equation}
\label{eqn:graph_Laplacian}
\Lap_{n,r} = \bf{D} - \bf{W}
\end{equation}
\emph{Laplacian eigenmaps} and \emph{Laplacian smoothing} each use $\Lap_{n,r}$ to construct an estimator $\wh{f}$ of $f_0$. Letting $v_1(G_{n,r}),\ldots,v_n(G_{n,r})$ denote eigenvectors of $\Lap_{n,r}$, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ \citep{belkin2003} is given by 
\begin{equation}
\label{eqn:laplacian_eigenmaps}
\wh{f}_{\textrm{LE}} = \sum_{k = 1}^{\kappa} \Dotp{Y}{v_k(G_{n,r})}_n v_{k}(G_{n,r})
\end{equation}
where $\kappa \in \{1,...,n\}$ is a tuning parameter. (Here $\dotp{u}{v}_n = \frac{1}{n}\sum_{i = 1}^{n} u_i v_i$ for vectors $u,v \in \Reals^n$.) The Laplacian smoothing estimator $\wh{f}_{\LS}$ \citep{smola2003} is given by
\begin{equation}
\label{eqn:laplacian_smoothing}
\wh{f}_{\LS}= \min_{f \in \Reals^n} \Bigl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T L^s f \Bigr\}
\end{equation}
where $\rho > 0$ acts a tuning parameter. Then, assuming~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are reasonable estimators of $f_0$, the size of the statistics
\begin{align}
T_{\LE} & = \norm{\wh{f}_{\LE}}_n^2 \label{eqn:laplacian_eigenmaps_test}\\
T_{\LS} & = \norm{\wh{f}_{\LS}}_n^2 \label{eqn:laplacian_smoothing_test}
\end{align}
can be used to test whether or not $f_0 = 0$.

\begin{itemize}
	\item \textbf{Motivate graph Laplacians.} There already exist a vast number of methods for both the nonparametric testing and, especially, the nonparametric estimation problems. Nevertheless, methods such as~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} deserve a place in the nonparametric regression toolbox. Their advantages include:
	\begin{itemize}
		\item Simplicity.
		\item Computational efficiency.
		\item Domain and density adaptivity.
		\item Practical success.
	\end{itemize}
	\item \textbf{Our contributions, short version.} For these reasons, there is a  substantial body of work assessing the statistical properties of graph Laplacian based methods. \textcolor{red}{(TODO): Include citations.} In this paper, we adopt a different perspective, and analyze~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} through the minimax lens. We show that for a variety of function classes $\mc{F}$, distributions $P$, and choices of domain $\mc{X}$ for which minimax rates are known, both~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are minimax optimal estimators (with loss measured in a suitable norm) and both~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} are minimax optimal test statistics. This lends additional theoretical support to the case for using graph Laplacian methods.
\end{itemize}

\section{Problem Setup and Background}
\label{sec:problem_setup_and_background}

Before we can state our main results, we must formally define minimaxity in both the estimation and testing problems. As we do so, we will remind the reader of the minimax optimal rates for some classical function classes $\mc{F}$ that we will henceforth consider. We will also briefly recall some (of the many) estimators and tests which achieve these rates; in particular we will focus on the \emph{truncated series} and \emph{smoothing spline} methods, which can be viewed as analogous to the graph Laplacian methods we consider.

\begin{itemize}
	\item \textbf{Minimax estimation.} For given integers $s > 0$ and $p > 0$, the Sobolev space $W^{s,p}(P)$ consists of all functions $f \in \Leb^p(P)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d)$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^p(P)$. The Sobolev $\{s,p\}$ norm is then 
	\begin{equation*}
	\norm{f}_{W^{s,p}(P)}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^p \,dP
	\end{equation*}
	and for a given $M > 0$, the corresponding ball is $W^{s,p}(P, M) = \set{f: \norm{f}_{W^{s,p}(P)} \leq M}$. In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space; we adopt the usual convention of writing $H^s(P) = W^{s,p}(P)$ and $H^s(P,M) = W^{s,2}(P,M)$. We will confine our attention to this special case hereafter. Let us also consider the following assumption

	\begin{enumerate}[(P1)]
		\item 
		\label{asmp:bounded_density}
		The distribution $P$ is defined on domain $\Xset = [0,1]^d$, and admits a density $p$ with respect to the uniform (Lebesgue) measure on $[0,1]^d$. The density $p$ is bounded away from $0$ and $\infty$,
		\begin{equation*}
		0 < p_{\min} < p(x) < p_{\max} < \infty
		\end{equation*}
		for all $x \in [0,1]^d$.
	\end{enumerate}
	The assumption~\ref{asmp:bounded_density} is a standard but significant assumption; we will postpone a more significant discussion of it until \textcolor{red}{a later point}. Assuming~\ref{asmp:bounded_density}, the minimax estimation rate over the Sobolev ellipsoids $H^{s}(P,M)$ is
	\begin{equation*}
	\inf_{\wh{f}} \sup_{f \in H^s(P,M)} \Ebb\Bigl[\norm{\wh{f} - f}_{L^2(P)}^2\Bigr] \asymp n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d$,}
	\end{equation*}
	where the infimum is over all measurable maps $\wh{f}(X_1,Y_1,\ldots,X_n,Y_n)$
	\item \textbf{Minimax hypothesis testing.}
	\item \textbf{Our contributions, long version.}
	\begin{itemize}
		\item \textbf{Minimax optimal estimation.}
		We show that with high probability,  $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2s/(2s + d)}$ whenever $f_0 \in C^s(\mc{X})$ for any $s \leq 1$, or $f_0 \in H^s(P)$ for $s = 1$. The same conclusions hold with respect to $\wh{f}_{\LS}$ when $d \leq 4$. 
		\item \textbf{Minimax optimal testing.}
		We show that with high probability, a level-$\alpha$ test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in C^s(\mc{X})$ for any $s \leq 1$---or $f_0 \in H^s(P)$ for $s = 1$---additionally satisfies $\norm{f_0}_{\Leb^2(P)} \gtrsim \max\{n^{-4s/(4s + d)},n^{-1/4}\}$. \textcolor{red}{The same conclusions hold with respect to $\wh{f}_{LS}$.}
		\item \textbf{Manifold adaptivity.}
		If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$, each of the aforementioned rates hold with $d$ replaced by $m$.
		\item \textbf{Taking advantage of higher-order smoothness conditions.} Under appropriate boundary conditions on the function classes $C^s$ and $H^s$ and additional smoothness conditions on $P$, when $s > 1$  each of the aforementioned conclusions hold for certain values of $d$.
	\end{itemize}
	\item \textbf{Organization.}
	\item \textbf{Notation.}
\end{itemize}

\section{Minimax optimal graph Laplacian methods}

The estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and the resulting test statistics~\eqref{eqn:laplacian_eigenmaps_test} and ~\eqref{eqn:laplacian_smoothing_test}, achieve the standard minimax optimal estimation and hypothesis testing rates, respectively.

\subsection{Minimax optimal estimation}

Under appropriate conditions, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves minimax rates over all $f_0 \in H_0^s(P,M)$ with high probability. These conditions will take two forms. The first pair of conditions will be on the neighborhood graph $G_{n,r}$. These assumptions are mild, as the construction of the graph is in the user's control.
\begin{enumerate}[label=(K\arabic*)]
	\item 
	\label{asmp:kernel_form}
	The kernel function $K(z) = \1\{\norm{z}_2/r \leq 1 \}$.
	\item 
	\label{asmp:kernel_radius}
	The radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
	\begin{equation*}
	n^{-1/(2(s - 1) + d)}\leq r(n) \leq M^{2/(2s + d)}n^{-1/(2s + d)}
	\end{equation*}
	for all $n$ sufficiently large.
\end{enumerate}

We will also need to make an additional assumption on $P$ (additional to~\ref{asmp:bounded_density}, that is). The assumption we will make is stronger than~\ref{asmp:bounded_density}, but we need it to make $\Lap_{{n,r}}$---or more precisely, the iterated Laplacian operator $\Lap_{{n,r}}^s$---behave \textcolor{red}{like an appropriate differential operator}.
\begin{enumerate}[label=(P\arabic*)]
	\setcounter{enumi}{1}
	\item 
	\label{asmp:smooth_density}The distribution $P$ satisfies~\ref{asmp:bounded_density}. Additionally, the density $p \in C^{\beta}([0,1]^d;p_{\max})$ for some $\beta \geq 0$. 
\end{enumerate}
Together these assumptions are used to ensure that the approximation error (bias) of $\wh{f}_{\LE}$ is of the correct order. Bounds on the estimation error of $\wh{f}_{\LE}$ -- both in expectation and in probability-- are straightforward. Together, bounds on these two sources of error imply Theorem~\ref{thm:laplacian_eigenmaps_estimation}.
\begin{theorem}
	\label{thm:laplacian_eigenmaps_estimation}
	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}. Suppose that $f_0 \in H_0^s(P,M)$, that $P$ satisfies~\ref{asmp:smooth_density} with $\beta = s - 1$, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius}. Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as in~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2s + d)}n^{d/(2s + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
	\begin{equation*}
	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4s + d)} n^{-2s/(2s + d)}
	\end{equation*}
	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
\end{theorem}

\begin{enumerate}
	\item \textcolor{red}{(TODO)}: Comment on the difference between this and minimax optimality. Cite others \textcolor{red}{(Wainwright, Liu)} who do the same thing.
	\item \textcolor{red}{(TODO)}: Comment on why we insist that $f_0$ satisfy a boundary condition.
	\item \textcolor{red}{(TODO)}: Comment a bit on the proof strategy; i.e., how we don't convergence properties of the eigenvectors $v_k$ towards eigenfunctions of the weighted Laplace-Beltrami operator, why we think that's advantageous, and how it might be limiting.
\end{enumerate}

\subsection{Minimax optimal hypothesis testing}

\section{Higher-order adaptivity}

\section{Fixed graph results}

\section{Simulations}
\begin{itemize}
	\item \textbf{Estimation MSE 1.} Plot $L^2(P_n)$ error as a function of $n$ and choice $f_0^{(n)}$ when $P$ is (close to) uniform, showing that Laplacian methods are competitive with their non-graph-based brethren.
	\item \textbf{Estimation fitted values.} Show fits for two regression problems, one where the distribution $P$ is assumed to be close to uniform, one where $P$ is dramatically non-uniform.
	\item \textbf{Estimation MSE 2.} Plot $L^2(P_n)$ error as a function of the distribution $P$, as we vary $P$ from being uniform to being very non-uniform.

	\item \textbf{Testing power.} 
\end{itemize}

\begin{figure}
	\centering
	\includegraphics[width = .48\textwidth]{plots/graph_regression/mean_squared_errors/sobolev_MSE_by_density}
	\caption{}
	\label{fig:piecewise_cosine}
\end{figure}

\clearpage

\appendix

Many of our results follow the same general two-part proof strategy. First, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to functionals defined on $G$. We then analyze the behavior of these functionals on the particular neighborhood graph $G_{n,r}$, giving high probability bounds which hold given our assumptions on the distribution $P$.

\section{Fixed graph error bounds}

\subsection{Laplacian eigenmaps on a fixed graph}
Suppose we observe $G = ([n],E)$ a fixed graph, and responses
\begin{equation}
\label{eqn:fixed_graph_regression_model}
Y_i = f_{0,i} + \varepsilon_i
\end{equation}
where $f_0 \in \Reals^n$ and $\varepsilon_i$ are independent $N(0,1)$ noise. The Laplacian eigenmaps estimator of $f_0$ on $G$ is then
\begin{equation}
\label{eqn:laplacian_eigenmaps_G}
\wh{f}_{\LE}(G) := \sum_{k = 1}^{\kappa} \biggl(\frac{1}{n}\sum_{i = 1}^{n} Y_i v_{k,i}(G) \biggr) v_k(G),
\end{equation}
where we note that \eqref{eqn:laplacian_eigenmaps} specializes~\eqref{eqn:laplacian_eigenmaps_G} to the case where $G = G_{n,r}$.

In Lemma~\ref{lem:fixed_graph_estimation}, we upper bound the mean squared error of $\wh{f}_{\LE}(G)$. Our bound will be stated as a function of $f_0^T \Lap_G^s f_0$--a measure of the smoothness the signal $f_0$ displays over the graph $G$--as well as the $\kappa$th eigenvalue $\lambda_{\kappa}(G)$.

\begin{lemma}
	\label{lem:fixed_graph_estimation}
	Suppose we observe a graph $G$ and responses $Y$ according to the model~\eqref{eqn:fixed_graph_regression_model}. Then for any $1 \leq \kappa \leq n$, 
	\begin{equation}
	\label{eqn:fixed_graph_estimation}
	\Ebb\Bigl[\norm{\wh{f}_{\LE}(G) - f_0}_n^2\Bigr] \leq 2\biggl(\frac{f_0^T \Lap_{G}^s f_0}{n \bigl(\lambda_{\kappa}(G)\bigr)^s} + \frac{\kappa}{n}\biggr)
	\end{equation}
	and
	\begin{equation}
	\label{eqn:fixed_graph_estimation_1}
	\norm{\wh{f}_{\LE}(G) - f_0}_n^2 \leq 2\biggl(\frac{f_0^T \Lap_{G}^s f_0}{n \bigl(\lambda_{\kappa}(G)\bigr)^s} + \frac{4\kappa}{n}\biggr)
	\end{equation}
	with probability at least $1 - \exp(-\kappa)$.
\end{lemma}
The two terms on the right hand side of~\eqref{eqn:fixed_graph_estimation} and~\eqref{eqn:fixed_graph_estimation_1} represent the squared bias and variance of the estimator $\wh{f}_{\LE}(G)$. Note that the squared bias--- and therefore the overall mean squared error---depends on the quadratic form $f_0^T \Lap_{G}^s f_0$ and the eigenvalue $\lambda_{\kappa}(G)$, whereas the variance depends instead on the dimension $\kappa$ of the subspace $V_{\kappa} = \textrm{span}\{v_1,\ldots,v_{\kappa}\}$ to which $\wh{f}_{\LE}(G)$ belongs. In this respect, the error bound on the Laplacian eigenmaps estimator resembles the error bound on the \textcolor{red}{orthogonal series} estimator (see e.g. \textcolor{red}{Tsybakov}).

To prove Theorems~\ref{thm:laplacian_eigenmaps_estimation},\textcolor{red}{...}, we would like to apply Lemmas~\ref{lem:fixed_graph_estimation},\textcolor{red}{...} in the random design context, i.e. to the random geometric graph $G_{n,r}$. To do so, we must establish the following (high probability) bounds: (i) an upper bound on the graph Sobolev seminorm of the regression function $f_0 \Lap_{n,r}^s f_0$, (ii) and lower boud on the Laplacian eigenvalue $\lambda_{\kappa}(G_{n,r})$ as a function of $\kappa$, and (iii) a lower bound on the empirical norm $\norm{f}_{n}$ as a function of $\norm{f}_{L^2(P)}$.

\section{Bounds on the graph Sobolev semi-norm}

\section{Bounds on graph eigenvalues}

\section{Proofs of Theorems}
\label{sec:proofs_theorems}

\subsection{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation}}
\label{subsec:laplacian_eigenmaps_estimation_pf}

\paragraph{\eqref{eqn:fixed_graph_estimation_sobolev_rate_2} holds with high probability.}

We first bound $f_0^T \Lap_{n,r}^s f_0$ in expectation.
\begin{lemma}
	\label{lem:roughness_functional_expectation_sobolev}
	Let $\Xset$ be a Lipschitz domain and let $s \geq 1$ be an integer. Suppose that \textcolor{red}{$f \in H_0^{s}(P,M)$}, and that $P$ admits a density $p \in C^{s-1}(\Xset,p_{\max})$ for some constant $p_{\max} < \infty$. Let $\{r(n)\}$ be a sequence of connectivity radii such that $n^{-1/(2(s - 1) + d)} \leq r(n)$ for all $n$ sufficiently large, and $r(n) \to 0$ as $n \to \infty$. Then \textcolor{red}{for any $2$nd-order kernel $K$}, there exists a constant $c$ depending only on $p_{\max}, K_{\max}$, $s$, $d$ and $\Xset$ such that the resulting neighborhood graph Laplacian satisfies
	\begin{equation}
	\label{eqn:roughness_functional_expectation_sobolev}
	\Ebb\bigl[f^T L_{n,r}^s f\bigr] \leq c \cdot n \cdot (nr^{d+2})^s \cdot M^2
	\end{equation}
	for all $n$ sufficiently large.
\end{lemma}

Since $\Lap_{n,r}$ is a positive semi-definite matrix, $f_0^T \Lap_{{n,r}}^s f_0$ must be non-negative. We may therefore apply Markov's inequality and conclude
\begin{equation}
\label{eqn:roughness_functional_probability_sobolev}
f_0^T \Lap_{{n,r}}^s f_0 \leq  \frac{1}{\delta} \cdot c \cdot n \cdot (nr^{d+2})^s \cdot M^2
\end{equation}
with probability at least $1 - \delta$.

\paragraph{\eqref{eqn:fixed_graph_estimation_sobolev_rate_1} holds with high probability.}

Lemma~\ref{lem:neighborhood_eigenvalue} establishes a scaling rate which holds for all eigenvalues $\lambda_{k}(G_{n,r})$, assuming appropriate regularity conditions on $P$.
\begin{lemma}
	\label{lem:neighborhood_eigenvalue}
	Suppose the distribution $P$ satisfies~\ref{asmp:bounded_density}. Then there exists constants $c$ and $C$ which depend only on $p_{\min}, p_{\max}$ and $d$, such that with probability at least $1 - C r^{-d} \exp\bigl\{- c n r^d\bigr\}$, for all $2 \leq k \leq n$
	\begin{equation}
	\label{eqn:neighborhood_eigenvalue}
	\lambda_k(G_{n,r}) \geq c \cdot \min\Bigl\{nr^{d + 2}k^{2/d}, nr^d\Bigr\}.
	\end{equation}	
\end{lemma}
The first term on the right hand side of~\eqref{eqn:neighborhood_eigenvalue} is the scaling of the eigenvalues we desire. The second term reflects the fact that the maximum eigenvalue of the Laplacian can never be more than twice the maximum degree of the graph. It is a bit of simple algebra to derive the conditions under which the first term is the smaller of the two when $k = \kappa = M^{2/(2s + d)} n^{d/(2s + d)}$.
\begin{corollary}
	\label{cor:neighborhood_eigenvalue}
	Let $\kappa = M^{2/(2s + d)} n^{d/(2s + d)}$. For any $r \leq M^{-2/(d(2s +d))} n^{-1/(2s + d)}$, 
	\begin{equation}
	\label{eqn:neighborhood_eigenvalue_1}
	\lambda_{\kappa}(G_{n,r}) \geq c \cdot nr^{d+2}{\kappa}^{2/d}
	\end{equation}
	with probability at least $1 - C r^{-d} \exp\{- c n r^d\}$. The constants $c$ and $C$ are the same as in Lemma~\ref{lem:neighborhood_eigenvalue}.
\end{corollary}

\paragraph{Putting the pieces together.}

Let $n$ be sufficiently large so that $n^{-1/(2(s - 1)+ d)} \leq r(n) \leq M^{-2/(d(2s + d))} n^{-1/(2s + d)}$. Then by~\eqref{eqn:roughness_functional_probability_sobolev} and~\eqref{eqn:neighborhood_eigenvalue_1}, there exists a set $\mathcal{E}(f_0) \subseteq \Xset^n$ of probability at least $1 - \delta - Cr^{-d}\exp\{-cnr^d\}$ such that whenever ${\bf X} \in \mathcal{E}(f_0)$,
\begin{equation*}
f_0^T \Lap_{n,r}^s f_0 \leq C \cdot n \cdot C_n^s \cdot M^2
\end{equation*}
and
\begin{equation*}
\lambda_{\kappa}(G_{n,r}) \geq c \cdot nr^{d+2}\kappa^{2/d}
\end{equation*}
Theorem~\ref{thm:laplacian_eigenmaps_estimation} then follows by Corollary~\ref{cor:fixed_graph_estimation_sobolev_rate}.

\bibliographystyle{plainnat}
\bibliography{../graph_testing_bibliography}

\end{document}

### 2020-08-05 notes/graph_regression2.tex
Added theorem on estimation rates for Laplacian smoothing.

@@
 where $\kappa \in \{1,...,n\}$ is a tuning parameter.
-(Here $\dotp{u}{v}_n = \frac{1}{n}\sum_{i = 1}^{n} u_i v_i$ for vectors $u,v \in \Reals^n$.) The Laplacian smoothing estimator $\wh{f}_{\LS}$ \citep{smola2003} is given by
+(Here $\dotp{u}{v}_n = \frac{1}{n}\sum_{i = 1}^{n} u_i v_i$ for vectors $u,v \in \Reals^n$.) The Laplacian smoothing estimator $\wt{f}_{\LS}$ \citep{smola2003} is given by
 \begin{equation}
 \label{eqn:laplacian_smoothing}
-\wh{f}_{\LS}= \min_{f \in \Reals^n} \Bigl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T L^s f \Bigr\}
+\wt{f}_{\LS}= \min_{f \in \Reals^n} \Bigl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T L^s f \Bigr\}
 \end{equation}
@@
 T_{\LE} & = \norm{\wh{f}_{\LE}}_n^2 \label{eqn:laplacian_eigenmaps_test}\\
-T_{\LS} & = \norm{\wh{f}_{\LS}}_n^2 \label{eqn:laplacian_smoothing_test}
+T_{\LS} & = \norm{\wt{f}_{\LS}}_n^2 \label{eqn:laplacian_smoothing_test}
 \end{align}
@@
 		We show that with high probability,  $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2s/(2s + d)}$ whenever $f_0 \in C^s(\mc{X})$ for any $s \leq 1$, or $f_0 \in H^s(P)$ for $s = 1$.
-The same conclusions hold with respect to $\wh{f}_{\LS}$ when $d \leq 4$.
+The same conclusions hold with respect to $\wt{f}_{\LS}$ when $d \leq 4$.
 		\item \textbf{Minimax optimal testing.}
@@
 
-The estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and the resulting test statistics~\eqref{eqn:laplacian_eigenmaps_test} and ~\eqref{eqn:laplacian_smoothing_test}, achieve the standard minimax optimal estimation and hypothesis testing rates, respectively.
+Let us start by assuming $f_0 \in H^1(P,M)$.
+In this case, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and the resulting test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test}, achieve the standard minimax optimal estimation and hypothesis testing rates, respectively.
 
@@
 
-Under appropriate conditions, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves minimax rates over all $f_0 \in H_0^s(P,M)$ with high probability.
+Under appropriate conditions, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves minimax rates over all $f_0 \in H^1(P,M)$ with high probability.
+The conditions we impose will be on the neighborhood graph $G_{n,r}$.
+As the construction of the graph is in the user's control, these assumptions are therefore mild.
+\begin{enumerate}[label=(K\arabic*)]
+	\item 
+	\label{asmp:kernel_form}
+	The kernel function $K(z) = \1\Bigl\{\norm{z}_2/r \leq 1 \Bigr\}$.
+	\item 
+	\label{asmp:kernel_radius}
+	The radii $r = r(n)$ satisfies $r(n) \to 0$ as $n \to \infty$, and additionally
+	\begin{equation*}
+	n^{-1/d}\leq r(n) \leq M^{2/(2 + d)}n^{-1/(2 + d)}
+	\end{equation*}
+	for all $n$ sufficiently large.
+\end{enumerate}
+
+We use~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius} to ensure that the approximation error (bias) of $\wh{f}_{\LE}$ is of the correct order.
+On the other hand, bounding the estimation error of $\wh{f}_{\LE}$ -- both in expectation and in probability-- is straightforward.
+Together, bounds on these two sources of error imply Theorem~\ref{thm:laplacian_eigenmaps_estimation1}.
+\begin{theorem}
+	\label{thm:laplacian_eigenmaps_estimation1}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
+Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius}.
+Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
+	\begin{equation*}
+	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4 + d)} n^{-2/(2 + d)}
+	\end{equation*}
+	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
+\end{theorem}
+
+\begin{itemize}
+	\item Discuss the choice of radius $r$ imposed by~\ref{asmp:kernel_radius}.
+	\item Comment on the distinction between error rates as measured by the norm in $L^2(P_n)$ vs. $L^2(P)$. 
+\end{itemize}
+
+Theorem~\ref{thm:laplacian_smoothing_estimation1} is analogous to Theorem~\ref{thm:laplacian_eigenmaps_estimation1}, dealing with the Laplacian smoothing estimator $\wt{f}_{\LS}$.
+The key difference is that it holds only when the dimension $d < 4$.
+\begin{theorem}
+	\label{thm:laplacian_smoothing_estimation1}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
+Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius}.
+Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
+	\begin{equation*}
+	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4 + d)} n^{-2/(2 + d)}
+	\end{equation*}
+	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
+\end{theorem}
+
+\begin{itemize}
+	\item A restriction on the maximum value of dimension $d$ such as $d \leq 4$ is unsurprising.
+In a related context (estimation over a $d$-dimensional lattice graph), \cite{sadhanala2016} require the same restriction.
+The underlying phenomenon is similar to that elucidated by the seminal papers~\cite{birge1993} and~\cite{birge1998}: when the dimension $d$ is too large, the entropy function is irregular and (penalized) least squares approaches are no longer minimax optimal.
+\end{itemize}
+
+\subsection{Minimax optimal hypothesis testing}
+
+\section{Higher-order smoothness classes.}
+\label{sec:higher_order_smoothness_classes}
+Suppose we are willing to assume $f_0$ possesses more than one smooth derivative.
+We can show that in this situation, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} obtain the faster (known) minimax rates in the nonparametric estimation and testing problems.
+However, these conclusions hold only for certain values of the dimension $d$.
+
+\subsection{Minimax optimal estimation}
+Under  conditions, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves minimax rates over all $f_0 \in H_0^s(P,M)$ with high probability.
 These conditions will take two forms.
@@
 \end{enumerate}
-
-\subsection{Minimax optimal hypothesis testing}
-
-\section{Higher-order adaptivity}
 

### 2020-08-09 notes/graph_regression2.tex
Added section on minimax testing with graph regression methods.

@@
 \begin{document}
-\title{Minimax-optimal learning with graph Laplacians}
+\title{Minimax-optimal nonparametric regression with graph Laplacians}
 \author{Alden Green}
@@
 \section{Introduction}
-
-\textcolor{red}{TODO:} Paragraph motivating graph Laplacian methods.
 
@@
 
-In graph based learning, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$, which should approximately capture the geometry of $P$ and $\mc{X}$, and then constructing an estimate $\wh{f}$ which is smooth with respect to $G_{n,r}$.
+In graph based nonparametric regression, the idea is to perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$.
 For a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$.
@@
 \end{equation}
-\emph{Laplacian eigenmaps} and \emph{Laplacian smoothing} each use $\Lap_{n,r}$ to construct an estimator $\wh{f}$ of $f_0$.
+The two graph-based regression methods we will consider are \emph{Laplacian eigenmaps} and \emph{Laplacian smoothing}.
+Each use $\Lap_{n,r}$ to construct an estimator $\wh{f}$ of $f_0$.
 Letting $v_1(G_{n,r}),\ldots,v_n(G_{n,r})$ denote eigenvectors of $\Lap_{n,r}$, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ \citep{belkin2003} is given by
@@
 where $\rho > 0$ acts a tuning parameter. Then, assuming~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are reasonable estimators of $f_0$, the size of the statistics
-\begin{align}
-T_{\LE} & = \norm{\wh{f}_{\LE}}_n^2 \label{eqn:laplacian_eigenmaps_test}\\
-T_{\LS} & = \norm{\wt{f}_{\LS}}_n^2 \label{eqn:laplacian_smoothing_test}
-\end{align}
+\begin{equation}
+T_{\LE} = \Bigl\|\wh{f}_{\LE}\Bigr\|_n^2 \label{eqn:laplacian_eigenmaps_test}
+\end{equation}
+and
+\begin{equation}
+T_{\LS} = \Bigl\|\wt{f}_{\LS}\Bigr\|_n^2 \label{eqn:laplacian_smoothing_test}
+\end{equation}
 can be used to test whether or not $f_0 = 0$.
@@
 This lends additional theoretical support to the case for using graph Laplacian methods.
+	\item \textbf{Summary}
 \end{itemize}
@@
 
-Before we can state our main results, we must formally define minimaxity in both the estimation and testing problems.
-As we do so, we will remind the reader of the minimax optimal rates for some classical function classes $\mc{F}$ that we will henceforth consider.
-We will also briefly recall some (of the many) estimators and tests which achieve these rates; in particular we will focus on the \emph{truncated series} and \emph{smoothing spline} methods, which can be viewed as analogous to the graph Laplacian methods we consider.
+Before we get to our main results, we will review some of the relevant literature on nonparametric estimation and regression, and recall some relevant minimax analysis.
+We will also briefly recall some (of the many) estimators and tests which achieve these rates.
+In particular, we will focus on the \emph{truncated series} and \emph{smoothing spline} methods.
+As we will see, Laplacian eigenmaps and Laplacian smoothing can be seen as graph-based analogues to these ``continuuum'' methods.
 
@@
 We will confine our attention to this special case hereafter.
-Let us also consider the following assumption
-
+Let us also make the following assumption:
 	\begin{enumerate}[(P1)]
@@
 	\end{enumerate}
-	The assumption~\ref{asmp:bounded_density} is a standard but significant assumption; we will postpone a more significant discussion of it until \textcolor{red}{a later point}.
+	The assumption~\ref{asmp:bounded_density} is a standard but significant assumption; we will return to discuss it further at \textcolor{red}{a later point}.
 Assuming~\ref{asmp:bounded_density}, the minimax estimation rate over the Sobolev ellipsoids $H^{s}(P,M)$ is
 	\begin{equation*}
-	\inf_{\wh{f}} \sup_{f \in H^s(P,M)} \Ebb\Bigl[\norm{\wh{f} - f}_{L^2(P)}^2\Bigr] \asymp n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d$,}
+	\inf_{\wh{f}} \sup_{f \in H^s(P,M)} \Ebb\Bigl[\norm{\wh{f} - f}_{L^2(P)}^2\Bigr] \asymp n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d \geq 0$,}
 	\end{equation*}
@@
 	\end{itemize}
-	\item \textbf{Organization.}
 	\item \textbf{Notation.}
@@
 \section{Minimax optimal graph Laplacian methods}
+\label{sec:minimax_optimal_graph_Laplacian_methods}
 
 Let us start by assuming $f_0 \in H^1(P,M)$.
-In this case, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and the resulting test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test}, achieve the standard minimax optimal estimation and hypothesis testing rates, respectively.
+In this section, we will establish that the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and the resulting test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test}, achieve the standard minimax optimal estimation and hypothesis testing rates, respectively.
 
@@
 	\item 
-	\label{asmp:kernel_radius}
+	\label{asmp:kernel_radius_estimation}
 	The radii $r = r(n)$ satisfies $r(n) \to 0$ as $n \to \infty$, and additionally
@@
 
-We use~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius} to ensure that the approximation error (bias) of $\wh{f}_{\LE}$ is of the correct order.
+We use~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_estimation} to ensure that the approximation error (bias) of $\wh{f}_{\LE}$ is of the correct order.
 On the other hand, bounding the estimation error of $\wh{f}_{\LE}$ -- both in expectation and in probability-- is straightforward.
@@
 	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
-Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius}.
+Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_estimation}.
 Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
@@
 \begin{itemize}
-	\item Discuss the choice of radius $r$ imposed by~\ref{asmp:kernel_radius}.
+	\item Discuss the choice of radius $r$ imposed by~\ref{asmp:kernel_radius_estimation}.
 	\item Comment on the distinction between error rates as measured by the norm in $L^2(P_n)$ vs. $L^2(P)$.
+	\item \textbf{Proof techniques}: to prove results we rely on both (a) classical bias-variance decomposition of orthogonal series estimators (see e.g.
+\textcolor{red}{Tsybakov}), and (b) pointwise and variational results on neighborhood graph Laplacian semi-norms and eigenvalues (see e.g.
+\textcolor{red}{Belkin04, Belkin08, GarciaTrillos18, Calder19})
 \end{itemize}
 
-Theorem~\ref{thm:laplacian_smoothing_estimation1} is analogous to Theorem~\ref{thm:laplacian_eigenmaps_estimation1}, dealing with the Laplacian smoothing estimator $\wt{f}_{\LS}$.
+Theorem~\ref{thm:laplacian_smoothing_estimation1} is analogous to Theorem~\ref{thm:laplacian_eigenmaps_estimation1}, except it deals with the Laplacian smoothing estimator $\wt{f}_{\LS}$ rather than the Laplacian eigenmaps estimator $\wh{f}_{\LE}$.
 The key difference is that it holds only when the dimension $d < 4$.
@@
 	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
-Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius}.
-Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
+Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_estimation}.
+Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when we set $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$ the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
 	\begin{equation*}
-	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4 + d)} n^{-2/(2 + d)}
+	\Bigl\|\wh{f}_{\LS} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4 + d)} n^{-2/(2 + d)}
 	\end{equation*}
@@
 \begin{itemize}
-	\item A restriction on the maximum value of dimension $d$ such as $d \leq 4$ is unsurprising.
+	\item \textbf{Proof techniques}: Because the solution to~\eqref{eqn:laplacian_smoothing} has an analytic form, we can rely on a bias-variance decomposition---as opposed to empirical process theory---to derive~Theorem~\ref{thm:laplacian_smoothing_estimation1}.
+	\item The restriction in Theorem~\ref{thm:laplacian_smoothing_estimation1} that the dimension $d$ be less than $4$ is unsurprising.
 In a related context (estimation over a $d$-dimensional lattice graph), \cite{sadhanala2016} require the same restriction.
-The underlying phenomenon is similar to that elucidated by the seminal papers~\cite{birge1993} and~\cite{birge1998}: when the dimension $d$ is too large, the entropy function is irregular and (penalized) least squares approaches are no longer minimax optimal.
+The underlying phenomenon is similar to that elucidated by the seminal papers~\cite{birge1993} and~\cite{birge1998}: when the dimension $d$ is at least $4$, the parameter space is too large, and (penalized) least squares approaches are no longer minimax optimal.
 \end{itemize}
@@
 
-\section{Higher-order smoothness classes.}
+Under appropriate conditions, the Laplacian eigenmaps test statistic $T_{\LE}$ has non-trivial power over all $f_0 \in H^1(P,M)$ that are separated from $0$ in $\Leb^2(P)$ norm by at least the minimax critical radius.
+These conditions include~\ref{asmp:kernel_form} and~\ref{asmp:smooth_density}, but not~\ref{asmp:kernel_radius_estimation}.
+Instead, we will require the following scaling of the neighborhood graph radius.
+\begin{enumerate}[label=(K\arabic*)]
+	\setcounter{enumi}{2}
+	\item 
+	\label{asmp:kernel_radius_testing}
+	The radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
+	\begin{equation*}
+	n^{-1/d} \leq r(n) \leq n^{-2/(4 + d)}
+	\end{equation*}
+	for all $n$ sufficiently large.
+\end{enumerate}
+
+\begin{theorem}
+	\label{thm:laplacian_eigenmaps_testing1}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}. Suppose that $f_0 \in H^1(P,M)$, that
+	$P$ satisfies~\ref{asmp:bounded_density}, and that the neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_testing}.
+	Then for the Laplacian eigenmaps test statistic $T_{\LE}$ as in~\eqref{eqn:laplacian_eigenmaps_test} with $\kappa = n^{2d/(4 + d)} M^{4d/(4 + d)}$, the following statement holds true for all $n$ sufficiently large: there exist constants $C_1$ and $C_2$ which do not depend on $f_0$ such that if
+	\begin{equation}
+	\label{eqn:laplacian_eigenmaps_testing}
+	\norm{f_0}_{\Leb^2(P)}^2 \geq \frac{C_1}{\delta^2} M^{2d/(4 + d)} n^{-4/(4 + d)}
+	\end{equation}
+	then
+	\begin{equation*}
+	\Pbb \biggl(T_{\LE} \geq \frac{\kappa}{n} + \frac{1}{\delta}\sqrt{\frac{2\kappa}{n^2}}\biggr) \geq 1 - C_2\Bigl(\delta + r^{-d} \exp\{- c n r^d\}\Bigr).
+	\end{equation*}
+\end{theorem}
+\begin{itemize}
+	\item Explain why Theorem~\ref{thm:laplacian_eigenmaps_testing1} implies minimax optimality.
+\end{itemize}
+
+The estimation and testing rates derived in Section~\ref{sec:minimax_optimal_graph_Laplacian_methods} exhibit the curse of dimensionality: as $d$ gets large, inference becomes hard.
+As covered in Section~\ref{sec:problem_setup_and_background}, it is well known that by making additional structural assumptions---either on the regression function $f_0$, the distribution $P$, or both---one can derive improved minimax rates.
+In the following two sections, we show that under such assumptions, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} can in certain situations obtain the faster (known) minimax rates in the nonparametric estimation and testing problems.
+
+\section{Higher-order smoothness classes}
 \label{sec:higher_order_smoothness_classes}
-Suppose we are willing to assume $f_0$ possesses more than one smooth derivative.
-We can show that in this situation, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} obtain the faster (known) minimax rates in the nonparametric estimation and testing problems.
-However, these conclusions hold only for certain values of the dimension $d$.
+In this section, we will assume that $f_0 \in H^s(P,M)$ for some integer $s > 1$. In other words, we will assume that $f_0$ has at least two smooth derivatives.
 
 \subsection{Minimax optimal estimation}
-Under  conditions, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves minimax rates over all $f_0 \in H_0^s(P,M)$ with high probability.
-These conditions will take two forms.
-The first pair of conditions will be on the neighborhood graph $G_{n,r}$.
-These assumptions are mild, as the construction of the graph is in the user's control.
+Under the following three conditions, the estimation error of the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves the minimax optimal rate for any $f_0 \in H^s(P,M)$.
 \begin{enumerate}[label=(K\arabic*)]
+	\setcounter{enumi}{2}
 	\item 
-	\label{asmp:kernel_form}
-	The kernel function $K(z) = \1\{\norm{z}_2/r \leq 1 \}$.
-	\item 
-	\label{asmp:kernel_radius}
-	The radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
+	\label{asmp:kernel_radius_higher_order}
+	\textbf{Neighborhood graph radius}:
+	The sequence of radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
 	\begin{equation*}
@@
 \end{enumerate}
-
-We will also need to make an additional assumption on $P$ (additional to~\ref{asmp:bounded_density}, that is).
-The assumption we will make is stronger than~\ref{asmp:bounded_density}, but we need it to make $\Lap_{{n,r}}$---or more precisely, the iterated Laplacian operator $\Lap_{{n,r}}^s$---behave \textcolor{red}{like an appropriate differential operator}.
+\begin{enumerate}[label=(F\arabic*)]
+	\setcounter{enumi}{0}
+	\item 
+	\label{asmp:boundary} \textbf{Boundary condition}; The regression function $f_0$ is zero-trace. In other words, $f_0 \in H_0^s(P,M)$. 
+\end{enumerate}
 \begin{enumerate}[label=(P\arabic*)]
@@
 	\item 
-	\label{asmp:smooth_density}The distribution $P$ satisfies~\ref{asmp:bounded_density}. Additionally, the density $p \in C^{\beta}([0,1]^d;p_{\max})$ for some $\beta \geq 0$. 
+	\label{asmp:smooth_density}\textbf{$\beta$-Holder density}: The design distribution $P$ has a density $p \in C^{\beta}([0,1]^d;p_{\max})$ for some $\beta \geq 0$. 
 \end{enumerate}
-Together these assumptions are used to ensure that the approximation error (bias) of $\wh{f}_{\LE}$ is of the correct order.
+
+We note that the assumption~\ref{asmp:kernel_radius_higher_order} is rather mild, since the construction of the neighborhood graph $G_{n,r}$ is in the user's control.
+whereas the latter two assumptions are more stringent.
+That being said, some restrictions on the boundary behavior of the regression function $f_0$ are typical in the analysis of orthogonal series estimators (see e.g.
+\textcolor{red}{Tsybakov}), as are assumptions on the smoothness of the density $p$ under a random design setup (see e.g.
+\textcolor{red}{Mukherjee16}).
+
+Together, conditions~\ref{asmp:boundary} and~\ref{asmp:smooth_density} are used to ensure that the approximation error (bias) of $\wh{f}_{\LE}$ is of the correct order.
 Bounds on the estimation error of $\wh{f}_{\LE}$ -- both in expectation and in probability-- are straightforward.
@@
 	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
-Suppose that $f_0 \in H_0^s(P,M)$, that $P$ satisfies~\ref{asmp:smooth_density} with $\beta = s - 1$, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius}.
+Suppose that $f_0 \in H_0^s(P,M)$, that $P$ satisfies~\ref{asmp:smooth_density} with $\beta = s - 1$, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_higher_order}.
 Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as in~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2s + d)}n^{d/(2s + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
@@
 
-\section{Fixed graph results}
+\section{Manifold Assumption}
+\label{sec:manifold_assumption}
+
+\section{Discussion}
+
+\begin{itemize}
+	\item Comment on the different types of relationships between $s$ and $d$.
+(1) Those due to fundamental lower bounds on testing; (2) those due to largeness of the parameter space for Laplacian smoothing, and (3) those which (potentially) could be improved by different proof techniques.
+\end{itemize}
 
@@
 Many of our results follow the same general two-part proof strategy.
-First, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to functionals defined on $G$.
-We then analyze the behavior of these functionals on the particular neighborhood graph $G_{n,r}$, giving high probability bounds which hold given our assumptions on the distribution $P$.
+First, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to functionals of the graph $G$.
+We then analyze the behavior of these functionals with respect to the particular neighborhood graph $G_{n,r}$ and giving high probability (upper or lower )bounds on these functionals.
+It is in this second step where we invoke our various assumptions on the distribution $P$ and regression function $f_0$.
 
 \section{Fixed graph error bounds}
+\label{sec:fixed_graph_error_bounds}
 
-\subsection{Laplacian eigenmaps on a fixed graph}
-Suppose we observe $G = ([n],E)$ a fixed graph, and responses
+Suppose we observe a graph $G = ([n],E)$, and responses
 \begin{equation}
@@
 \begin{equation}
-\label{eqn:laplacian_eigenmaps_G}
+\label{eqn:le_G}
 \wh{f}_{\LE}(G) := \sum_{k = 1}^{\kappa} \biggl(\frac{1}{n}\sum_{i = 1}^{n} Y_i v_{k,i}(G) \biggr) v_k(G),
 \end{equation}
-where we note that \eqref{eqn:laplacian_eigenmaps} specializes~\eqref{eqn:laplacian_eigenmaps_G} to the case where $G = G_{n,r}$.
+the Laplacian smoothing estimator of $f_0$ on $G$ is
+\begin{equation}
+\label{eqn:ls_G}
+\wh{f}_{\LS}(G) = \argmin_{f \in \Reals^n} \biggl\{ \sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_G^{s}  f \biggr\} = (\rho \Lap_G^s + I)^{-1}Y
+\end{equation}
+and their associated test statistics are
+\begin{equation}
+\label{eqn:le_ts_G}
+T_{\LE}(G) := \Bigl\|\wh{f}_{\LE}(G)\Bigr\|_n^2
+\end{equation}
+and
+\begin{equation}
+\label{eqn:ls_ts_G}
+T_{\LS}(G) := \Bigl\|\wt{f}_{\LS}(G)\Bigr\|_n^2
+\end{equation}
+We note that~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} specialize~\eqref{eqn:le_G}-\eqref{eqn:ls_ts_G} to the case where $G$ = $G_{n,r}$.
 
-In Lemma~\ref{lem:fixed_graph_estimation}, we upper bound the mean squared error of $\wh{f}_{\LE}(G)$.
+\subsection{Estimation: Laplacian eigenmaps on a fixed graph}
+
+In Lemma~\ref{lem:le_fixed_graph_estimation}, we upper bound the mean squared error of $\wh{f}_{\LE}(G)$.
 Our bound will be stated as a function of $f_0^T \Lap_G^s f_0$--a measure of the smoothness the signal $f_0$ displays over the graph $G$--as well as the $\kappa$th eigenvalue $\lambda_{\kappa}(G)$.
@@
 \begin{lemma}
-	\label{lem:fixed_graph_estimation}
-	Suppose we observe a graph $G$ and responses $Y$ according to the model~\eqref{eqn:fixed_graph_regression_model}. Then for any $1 \leq \kappa \leq n$, 
+	\label{lem:le_fixed_graph_estimation}
+	Let $G$ be a fixed (non-random) graph. Suppose we observe $G$ and responses $Y$ according to the model~\eqref{eqn:fixed_graph_regression_model}. Then for any $1 \leq \kappa \leq n$, 
 	\begin{equation}
-	\label{eqn:fixed_graph_estimation}
-	\Ebb\Bigl[\norm{\wh{f}_{\LE}(G) - f_0}_n^2\Bigr] \leq 2\biggl(\frac{f_0^T \Lap_{G}^s f_0}{n \bigl(\lambda_{\kappa}(G)\bigr)^s} + \frac{\kappa}{n}\biggr)
+	\label{eqn:le_fixed_graph_estimation}
+	\Ebb\Bigl[\norm{\wh{f}_{\LE}(G) - f_0}_n^2\Bigr] \leq 2\biggl(\frac{f_0^T \Lap_{G}^s f_0}{n \bigl(\lambda_{\kappa}(G)\bigr)^s} + \frac{\kappa}{n}\biggr).
 	\end{equation}
-	and
+	Additionally, 
 	\begin{equation}
-	\label{eqn:fixed_graph_estimation_1}
+	\label{eqn:le_fixed_graph_estimation_1}
 	\norm{\wh{f}_{\LE}(G) - f_0}_n^2 \leq 2\biggl(\frac{f_0^T \Lap_{G}^s f_0}{n \bigl(\lambda_{\kappa}(G)\bigr)^s} + \frac{4\kappa}{n}\biggr)
@@
 \end{lemma}
-The two terms on the right hand side of~\eqref{eqn:fixed_graph_estimation} and~\eqref{eqn:fixed_graph_estimation_1} represent the squared bias and variance of the estimator $\wh{f}_{\LE}(G)$.
+The two terms on the right hand side of~\eqref{eqn:le_fixed_graph_estimation} and~\eqref{eqn:le_fixed_graph_estimation_1} represent the squared bias and variance of the estimator $\wh{f}_{\LE}(G)$.
 Note that the squared bias--- and therefore the overall mean squared error---depends on the quadratic form $f_0^T \Lap_{G}^s f_0$ and the eigenvalue $\lambda_{\kappa}(G)$, whereas the variance depends instead on the dimension $\kappa$ of the subspace $V_{\kappa} = \textrm{span}\{v_1,\ldots,v_{\kappa}\}$ to which $\wh{f}_{\LE}(G)$ belongs.
@@
 
+\paragraph{Proof of Lemma~\ref{lem:le_fixed_graph_estimation}:}
+
+\textcolor{red}{(TODO)}
+
+\subsection{Estimation: Laplacian smoothing on a fixed graph.}
+
+\begin{lemma}
+	\label{lem:ls_fixed_graph_estimation}
+	Suppose we observe a graph $G$ and responses $Y$ according to the model~\eqref{eqn:fixed_graph_regression_model}. Then for any $\rho > 0$ and positive integer $s$,
+	\begin{equation}
+	\label{eqn:ls_fixed_graph_estimation}
+	\Ebb\Bigl[\norm{\wt{f}_{\LS}(G) - f_0}_n^2\Bigr] \leq 2\biggl(\frac{\rho \cdot f_0^T \Lap_{G}^s f_0}{n} + \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{(\rho \cdot \lambda_{k}(G)^s + 1)^2}\biggr)
+	\end{equation}
+	Additionally,
+	\begin{equation}
+	\label{eqn:ls_fixed_graph_estimation_prob}
+	\Pbb\Biggl[\norm{\wt{f}_{\LS}(G) - f_0}_n^2 \leq 2\biggl(\frac{\rho \cdot f_0^T \Lap_{G}^s f_0}{n} + \frac{4}{n}\sum_{k = 1}^{n} \frac{1}{(\rho \cdot \lambda_{k}(G)^s + 1)^2}\biggr)\Biggr] \geq 1 - \exp\biggl(-\sum_{k = 1}^{n}\frac{1}{(\rho \lambda_k(G)^s + 1)^2}\biggr)
+	\end{equation}
+\end{lemma}
+
+\paragraph{Proof of Lemma~\ref{lem:ls_fixed_graph_estimation}:}
+
+\textcolor{red}{(TODO)}
+
+\subsection{Testing: Laplacian eigenmaps on a fixed graph.}
+
+\begin{lemma}
+	\label{lem:fixed_graph_testing}
+	Suppose we observe a graph $G$ and responses $Y$ according to the model~\eqref{eqn:fixed_graph_regression_model}. Then for any $1 \leq \kappa \leq n$, each of the following holds for any $b \geq 1$,
+	\begin{enumerate}
+		\item \textbf{Type I error:} Under the null hypothesis $f_0 = 0$,
+		\begin{equation}
+		\label{eqn:graph_spectral_type_I_error}
+		\Pbb\biggl(T_{\LE}(G) \geq \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}\biggr) \leq \frac{2}{b^2}.
+		\end{equation}
+		\item \textbf{Type II error:} Assume that
+		\begin{equation}
+		\label{eqn:fixed_graph_testing_critical_radius_1}
+		\Bigl\|\sum_{k = 1}^{\kappa} \dotp{f_0}{n}_n v_k\Bigr\|_n^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}}.
+		\end{equation}
+		Then,
+		\begin{equation}
+		\label{eqn:graph_spectral_type_II_error}
+		\Pbb\biggl(T_{\LE}(G) \leq \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}\biggr) \leq \frac{2}{b^2} + \frac{8}{b\sqrt{2\kappa}}.
+		\end{equation}
+		In particular if
+		\begin{equation}
+		\label{eqn:fixed_graph_testing_critical_radius}
+		\frac{1}{n} \sum_{i = 1}^{n} f_{0,i}^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{f_0^T L_G^s f_0}{n\lambda_{\kappa}(G)^s}
+		\end{equation}
+		then~\eqref{eqn:fixed_graph_testing_critical_radius_1} and thus~\eqref{eqn:graph_spectral_type_II_error} follow.
+	\end{enumerate}
+\end{lemma}
+
+\subsection{Testing: Laplacian smoothing on a fixed graph.}
+
+\clearpage
 To prove Theorems~\ref{thm:laplacian_eigenmaps_estimation},\textcolor{red}{...}, we would like to apply Lemmas~\ref{lem:fixed_graph_estimation},\textcolor{red}{...} in the random design context, i.e.
@@
 \section{Bounds on the graph Sobolev semi-norm}
-
-\section{Bounds on graph eigenvalues}
-
-\section{Proofs of Theorems}
-\label{sec:proofs_theorems}
-
-\subsection{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation}}
-\label{subsec:laplacian_eigenmaps_estimation_pf}
-
-\paragraph{\eqref{eqn:fixed_graph_estimation_sobolev_rate_2} holds with high probability.}
-
+\label{sec:graph_sobolev_seminorm}
 We first bound $f_0^T \Lap_{n,r}^s f_0$ in expectation.
@@
 
-Since $\Lap_{n,r}$ is a positive semi-definite matrix, $f_0^T \Lap_{{n,r}}^s f_0$ must be non-negative. We may therefore apply Markov's inequality and conclude
-\begin{equation}
-\label{eqn:roughness_functional_probability_sobolev}
-f_0^T \Lap_{{n,r}}^s f_0 \leq  \frac{1}{\delta} \cdot c \cdot n \cdot (nr^{d+2})^s \cdot M^2
-\end{equation}
-with probability at least $1 - \delta$.
-
-\paragraph{\eqref{eqn:fixed_graph_estimation_sobolev_rate_1} holds with high probability.}
-
+\section{Bounds on graph eigenvalues}
+\label{sec:graph_eigenvalues}
 Lemma~\ref{lem:neighborhood_eigenvalue} establishes a scaling rate which holds for all eigenvalues $\lambda_{k}(G_{n,r})$, assuming appropriate regularity conditions on $P$.
@@
 
+\section{Proofs of Theorems}
+\label{sec:proofs_theorems}
+With the results of Sections~\ref{sec:fixed_graph_error_bounds}-\ref{sec:graph_eigenvalues} established, the proofs of our main theorems are quite straightforward.
+
+\subsection{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation}}
+\label{subsec:laplacian_eigenmaps_estimation_pf}
+
+\paragraph{\eqref{eqn:fixed_graph_estimation_sobolev_rate_2} holds with high probability.}
+
+Since $\Lap_{n,r}$ is a positive semi-definite matrix, $f_0^T \Lap_{{n,r}}^s f_0$ must be non-negative. We may therefore apply Markov's inequality and conclude
+\begin{equation}
+\label{eqn:roughness_functional_probability_sobolev}
+f_0^T \Lap_{{n,r}}^s f_0 \leq  \frac{1}{\delta} \cdot c \cdot n \cdot (nr^{d+2})^s \cdot M^2
+\end{equation}
+with probability at least $1 - \delta$.
+
+\paragraph{\eqref{eqn:fixed_graph_estimation_sobolev_rate_1} holds with high probability.}
+
 \paragraph{Putting the pieces together.}
@@
 
+\section{Concentration Inequalities}
+
 \bibliographystyle{plainnat}

### 2020-08-12 notes/graph_regression2.tex

@@
 
-In the random design nonparametric regression setup, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ on a domain $\Xset \subset \Rd$ and
+In the random design nonparametric regression setup, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Rd$, and
 \begin{equation}
@@
 with independent Gaussian noise $\varepsilon_i \sim N(0,1)$.
-The goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether or not $f_0 = 0$, i.e whether or not there is a signal present.
+Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether or not $f_0 = 0$, i.e whether or not there is a signal present.
 
@@
 \label{eqn:laplacian_smoothing}
-\wt{f}_{\LS}= \min_{f \in \Reals^n} \Bigl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T L^s f \Bigr\}
+\wt{f}_{\LS}= \min_{f \in \Reals^n} \Bigl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_{n,r}^s f \Bigr\}
 \end{equation}
@@
 	\item \textbf{Motivate graph Laplacians.} There already exist a vast number of methods for both the nonparametric testing and, especially, the nonparametric estimation problems.
+In fact, as we will see in Section~\ref{sec:problem_setup_and_background} some of these methods are actually quite analogous to the graph Laplacian estimation and testing methods we have just introduced.
 Nevertheless, methods such as~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} deserve a place in the nonparametric regression toolbox.
@@
 Theorem~\ref{thm:laplacian_smoothing_estimation1} is analogous to Theorem~\ref{thm:laplacian_eigenmaps_estimation1}, except it deals with the Laplacian smoothing estimator $\wt{f}_{\LS}$ rather than the Laplacian eigenmaps estimator $\wh{f}_{\LE}$.
-The key difference is that it holds only when the dimension $d < 4$.
 \begin{theorem}
@@
 	\item \textbf{Proof techniques}: Because the solution to~\eqref{eqn:laplacian_smoothing} has an analytic form, we can rely on a bias-variance decomposition---as opposed to empirical process theory---to derive~Theorem~\ref{thm:laplacian_smoothing_estimation1}.
-	\item The restriction in Theorem~\ref{thm:laplacian_smoothing_estimation1} that the dimension $d$ be less than $4$ is unsurprising.
-In a related context (estimation over a $d$-dimensional lattice graph), \cite{sadhanala2016} require the same restriction.
+	\item A key difference between Theorems~\ref{thm:laplacian_eigenmaps_estimation1} and~\ref{thm:laplacian_smoothing_estimation1} is that the latter holds only when $d \leq 4$.
+This restriction is unsurprising, and reflects a fundamental drawback in (penalized) least squares estimators more generally.
 The underlying phenomenon is similar to that elucidated by the seminal papers~\cite{birge1993} and~\cite{birge1998}: when the dimension $d$ is at least $4$, the parameter space is too large, and (penalized) least squares approaches are no longer minimax optimal.
+We should also mention that in a related context (estimation over a $d$-dimensional lattice graph), \cite{sadhanala2016} require the same restriction.
 \end{itemize}
@@
 	\label{thm:laplacian_eigenmaps_testing1}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}. Suppose that $f_0 \in H^1(P,M)$, that
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d \leq 3$. Suppose that $f_0 \in H^1(P,M)$, that
 	$P$ satisfies~\ref{asmp:bounded_density}, and that the neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_testing}.
@@
 	\item Explain why Theorem~\ref{thm:laplacian_eigenmaps_testing1} implies minimax optimality.
+	\item Include results regarding the low-smoothness regime (i.e. the case when $d \geq 4$.)
 \end{itemize}
 
-The estimation and testing rates derived in Section~\ref{sec:minimax_optimal_graph_Laplacian_methods} exhibit the curse of dimensionality: as $d$ gets large, inference becomes hard.
+\section{Higher-order smoothness classes}
+\label{sec:higher_order_smoothness_classes}
+The estimation and testing rates derived in Section~\ref{sec:minimax_optimal_graph_Laplacian_methods} exhibit a classic curse of dimensionality: as $d$ gets larger, inference becomes (much) harder.
 As covered in Section~\ref{sec:problem_setup_and_background}, it is well known that by making additional structural assumptions---either on the regression function $f_0$, the distribution $P$, or both---one can derive improved minimax rates.
@@
 
-\section{Higher-order smoothness classes}
-\label{sec:higher_order_smoothness_classes}
 In this section, we will assume that $f_0 \in H^s(P,M)$ for some integer $s > 1$. In other words, we will assume that $f_0$ has at least two smooth derivatives.
@@
 \subsection{Minimax optimal estimation}
+\label{subsec:minimax_estimation_higher_order}
 Under the following three conditions, the estimation error of the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves the minimax optimal rate for any $f_0 \in H^s(P,M)$.
@@
 
-We note that the assumption~\ref{asmp:kernel_radius_higher_order} is rather mild, since the construction of the neighborhood graph $G_{n,r}$ is in the user's control.
-whereas the latter two assumptions are more stringent.
-That being said, some restrictions on the boundary behavior of the regression function $f_0$ are typical in the analysis of orthogonal series estimators (see e.g.
-\textcolor{red}{Tsybakov}), as are assumptions on the smoothness of the density $p$ under a random design setup (see e.g.
-\textcolor{red}{Mukherjee16}).
-
-Together, conditions~\ref{asmp:boundary} and~\ref{asmp:smooth_density} are used to ensure that the approximation error (bias) of $\wh{f}_{\LE}$ is of the correct order.
-Bounds on the estimation error of $\wh{f}_{\LE}$ -- both in expectation and in probability-- are straightforward.
-Together, bounds on these two sources of error imply Theorem~\ref{thm:laplacian_eigenmaps_estimation}.
 \begin{theorem}
@@
 
+We note that the assumption~\ref{asmp:kernel_radius_higher_order} is rather mild, since the user can choose how to construct the neighborhood graph $G_{n,r}$.
+The assumptions~\ref{asmp:boundary} and~\ref{asmp:smooth_density} are more stringent.
+That being said, some restrictions on the boundary behavior of the regression function $f_0$ are typical in the analysis of orthogonal series estimators (see e.g.
+\textcolor{red}{Tsybakov}).
+On the other hand, restrictions on the smoothness of the density $p$ are also typical when analyzing orthogonal series estimators under a random design setup (see e.g.
+\textcolor{red}{Mukherjee16}).
+
 \begin{enumerate}
-	\item \textcolor{red}{(TODO)}: Comment on the difference between this and minimax optimality. Cite others \textcolor{red}{(Wainwright, Liu)} who do the same thing.
-	\item \textcolor{red}{(TODO)}: Comment on why we insist that $f_0$ satisfy a boundary condition.
-	\item \textcolor{red}{(TODO)}: Comment a bit on the proof strategy; i.e., how we don't convergence properties of the eigenvectors $v_k$ towards eigenfunctions of the weighted Laplace-Beltrami operator, why we think that's advantageous, and how it might be limiting.
+	\item \textcolor{red}{(TODO)}: Comment a bit on the proof strategy; i.e., 
 \end{enumerate}
+
+\subsection{Minimax optimal hypothesis testing}
+\label{subsec:minimax_testing_higher_order}
+
+Under appropriate conditions, the Laplacian eigenmaps test statistic $T_{\LE}$ has non-trivial power over all $f_0 \in H_0^s(P,M)$ that are separated from $0$ in $\Leb^2(P)$ norm by at least the minimax critical radius.
+These conditions include~\ref{asmp:kernel_form} and~\ref{asmp:smooth_density}.
+Additionally, we will require the following scaling of the neighborhood graph radii.
+\begin{enumerate}[label=(K\arabic*)]
+	\setcounter{enumi}{2}
+	\item 
+	\label{asmp:kernel_radius_testing}
+	The radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
+	\begin{equation*}
+	n^{-1/(2(s - 1) + d)} \leq r(n)~~\textrm{and}~~ r(n) \leq n^{-2/(4s + d)}
+	\end{equation*}
+	for all $n$ sufficiently large.
+\end{enumerate}
+In one sense this is still a mild condition, in that the neighborhood graph radius is still a tuning parameter to be selected by the user.
+However, some elementary calculations show that when $d > 4$, it is in fact the case that $n^{-1/(2(s - 1) + d)} > n^{-2/(4s + d)}$; hence no choice of $r$ can satisfy~\ref{asmp:kernel_radius_testing}.
+As a result, our theory applies only when $d \leq 4$.
+
+\begin{theorem}
+	\label{thm:laplacian_eigenmaps_testing}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}. Suppose that $d \leq 4$, that $f_0 \in H_0^s(P,M)$, that
+	$P$ satisfies~\ref{asmp:smooth_density} with $\beta = s - 1$, and that the neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_testing}.
+	Then for the Laplacian eigenmaps test statistic $T_{\LE}$ given by~\eqref{eqn:laplacian_eigenmaps_test} with tuning parameter $\kappa = n^{2d/(4s + d)} M^{4d/(4s + d)}$, the following statement holds true for all $n$ sufficiently large: there exist constants $C_1$ and $C_2$ which do not depend on $f_0$ such that if
+	\begin{equation}
+	\label{eqn:laplacian_eigenmaps_testing}
+	\norm{f_0}_{\Leb^2(P)}^2 \geq \frac{C_1}{\delta^2} M^{2d/(4s + d)} n^{-4s/(4s + d)}
+	\end{equation}
+	then
+	\begin{equation*}
+	\Pbb \biggl(T_{\LE} \geq \frac{\kappa}{n} + \frac{1}{\delta}\sqrt{\frac{2\kappa}{n^2}}\biggr) \geq 1 - C_2\Bigl(\delta + r^{-d} \exp\{- c n r^d\}\Bigr).
+	\end{equation*}
+\end{theorem}
+
+\begin{itemize}
+	\item \textbf{Test based on the Laplacian smoothing test statistic}. 
+	\item \textbf{Low-smoothness regime}. Include results on minimax testing when $4s < d$.
+	\item \textbf{Head off potential criticism}.
+Acknowledge frankly that there is a regime---when $4 < d < 4s$---where minimax testing rates are known, but for which we cannot prove that either of our test statistics achieve these rates.
+\end{itemize}
 
@@
 \begin{itemize}
-	\item Comment on the different types of relationships between $s$ and $d$.
-(1) Those due to fundamental lower bounds on testing; (2) those due to largeness of the parameter space for Laplacian smoothing, and (3) those which (potentially) could be improved by different proof techniques.
+	\item \textbf{Relationships between $s$ and $d$}.
+Comment on the different types of relationships between $s$ and $d$.
+(1) Those due to fundamental lower bounds on testing; (2) those due to ``defects'' in the proposed methods, i.e.
+largeness of the parameter space for Laplacian smoothing, and (3) those (potentially) due to ``defects'' in our proof strategy.
+	\item \textbf{Proof techniques}.
+On the subject of our proof techniques, the careful reader will note that our proofs do not rely on convergence of the Laplacian eigenvectors $v_k(G_{n,r})$ towards eigenfunctions of a limiting (weighted) Laplace-Beltrami operator.
+
+	\textcolor{red}{(TODO) Flesh the following out.} This is helpful for at least the following two reasons.
+First, proving convergence of eigenvectors is harder than proving convergence of eigenvalues, a fact which is reflected in the gap between the best known rates for the two problems.
+Second, because in order for our methods to be minimax optimal, orthogonal series estimators rely on projecting onto $\kappa$ basis functions, where $\kappa$ is a number growing in $n$.
+Thus we have an accumulation of error problem.
+	\item \textbf{Possible extensions}.
 \end{itemize}
@@
 		\end{equation}
-		\item \textbf{Type II error:} Assume that
+		\item \textbf{Type II error:} Under the alternative hypothesis $f_0 \neq 0$, suppose that
 		\begin{equation}
 		\label{eqn:fixed_graph_testing_critical_radius_1}
-		\Bigl\|\sum_{k = 1}^{\kappa} \dotp{f_0}{n}_n v_k\Bigr\|_n^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}}.
+		\Bigl\|\sum_{k = 1}^{\kappa} \dotp{f_0}{v_k}_n v_k\Bigr\|_n^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}}.
 		\end{equation}
@@
 		\label{eqn:fixed_graph_testing_critical_radius}
-		\frac{1}{n} \sum_{i = 1}^{n} f_{0,i}^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{f_0^T L_G^s f_0}{n\lambda_{\kappa}(G)^s}
+		\frac{1}{n} \sum_{i = 1}^{n} (f_{0,i})^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{f_0^T L_G^s f_0}{n\lambda_{\kappa}(G)^s}
 		\end{equation}
@@
 \end{lemma}
-The first term on the right hand side of~\eqref{eqn:neighborhood_eigenvalue} is the scaling of the eigenvalues we desire.
-The second term reflects the fact that the maximum eigenvalue of the Laplacian can never be more than twice the maximum degree of the graph.
-It is a bit of simple algebra to derive the conditions under which the first term is the smaller of the two when $k = \kappa = M^{2/(2s + d)} n^{d/(2s + d)}$.
-\begin{corollary}
-	\label{cor:neighborhood_eigenvalue}
-	Let $\kappa = M^{2/(2s + d)} n^{d/(2s + d)}$. For any $r \leq M^{-2/(d(2s +d))} n^{-1/(2s + d)}$, 
+As those readers familiar with Sobolev spaces will be aware of, the scaling $k^{2/d}$ is the \textcolor{red}{typical scaling of a Sobolev ellipsoid}, and is thus the scaling of the eigenvalues we desire.
+On the other hand, the second term on the right hand side of~\eqref{eqn:neighborhood_eigenvalue} is an artifact of the well known inequality $\lambda_n(G) \leq 2 \max_{i = 1,\ldots,n} \mathbf{D}_{ii}$--- in words, that the maximum eigenvalue of the Laplacian can never be more than twice the maximum degree of the graph.
+Doing some algebra, we may deduce the conditions on $k$ and $r$ under which the graph eigenvalue follows the desired scaling, i.e.
+$\lambda_k(G_{n,r}) \sim n r^{d + 2}k^{2/d}$.
+
+First, when $k = M^{2/(2s + d)} n^{d/(2s + d)}$, it holds that for any $r \leq M^{-2/(d(2s +d))} n^{-1/(2s + d)}$, 
 \begin{equation}
 \label{eqn:neighborhood_eigenvalue_1}
-	\lambda_{\kappa}(G_{n,r}) \geq c \cdot nr^{d+2}{\kappa}^{2/d}
+\lambda_{k}(G_{n,r}) \geq c \cdot nr^{d+2}{k}^{2/d}
 \end{equation}
-	with probability at least $1 - C r^{-d} \exp\{- c n r^d\}$. The constants $c$ and $C$ are the same as in Lemma~\ref{lem:neighborhood_eigenvalue}.
-\end{corollary}
+with probability at least $1 - C r^{-d} \exp\{- c n r^d\}$. Second, when $k$
 

### 2020-08-12 notes/weekly_notes/8.12.20.NotesForWeek.tex
New week, new weekly notes document.

--- 2020-08-05 notes/graph_regression2.tex
@@
 \begin{document}
-\title{Minimax-optimal learning with graph Laplacians}
+\title{Notes for Week 8/12/2020 - 8/19/2020}
 \author{Alden Green}
@@
 
-\section{Introduction}
-
-\textcolor{red}{TODO:} Paragraph motivating graph Laplacian methods.
-
-In the random design nonparametric regression setup, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ on a domain $\Xset \subset \Rd$ and
-\begin{equation}
-\label{eqn:random_design_regression}
-Y_i = f_0(X_i) + \varepsilon_i
-\end{equation}
-with independent Gaussian noise $\varepsilon_i \sim N(0,1)$.
-The goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether or not $f_0 = 0$, i.e whether or not there is a signal present.
-
-In graph based learning, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$, which should approximately capture the geometry of $P$ and $\mc{X}$, and then constructing an estimate $\wh{f}$ which is smooth with respect to $G_{n,r}$.
-For a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$.
-The weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ is an $n \times n$ matrix with entries
-\begin{equation*}
-\label{eqn:neighborhood_graph}
-{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr),
-\end{equation*}
-and the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$.
-Our star operator, the neighborhood graph Laplacian, can then be written as
-\begin{equation}
-\label{eqn:graph_Laplacian}
-\Lap_{n,r} = \bf{D} - \bf{W}
-\end{equation}
-\emph{Laplacian eigenmaps} and \emph{Laplacian smoothing} each use $\Lap_{n,r}$ to construct an estimator $\wh{f}$ of $f_0$.
-Letting $v_1(G_{n,r}),\ldots,v_n(G_{n,r})$ denote eigenvectors of $\Lap_{n,r}$, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ \citep{belkin2003} is given by
-\begin{equation}
-\label{eqn:laplacian_eigenmaps}
-\wh{f}_{\textrm{LE}} = \sum_{k = 1}^{\kappa} \Dotp{Y}{v_k(G_{n,r})}_n v_{k}(G_{n,r})
-\end{equation}
-where $\kappa \in \{1,...,n\}$ is a tuning parameter.
-(Here $\dotp{u}{v}_n = \frac{1}{n}\sum_{i = 1}^{n} u_i v_i$ for vectors $u,v \in \Reals^n$.) The Laplacian smoothing estimator $\wt{f}_{\LS}$ \citep{smola2003} is given by
-\begin{equation}
-\label{eqn:laplacian_smoothing}
-\wt{f}_{\LS}= \min_{f \in \Reals^n} \Bigl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T L^s f \Bigr\}
-\end{equation}
-where $\rho > 0$ acts a tuning parameter. Then, assuming~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are reasonable estimators of $f_0$, the size of the statistics
-\begin{align}
-T_{\LE} & = \norm{\wh{f}_{\LE}}_n^2 \label{eqn:laplacian_eigenmaps_test}\\
-T_{\LS} & = \norm{\wt{f}_{\LS}}_n^2 \label{eqn:laplacian_smoothing_test}
-\end{align}
-can be used to test whether or not $f_0 = 0$.
-
-\begin{itemize}
-	\item \textbf{Motivate graph Laplacians.} There already exist a vast number of methods for both the nonparametric testing and, especially, the nonparametric estimation problems.
-Nevertheless, methods such as~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} deserve a place in the nonparametric regression toolbox.
-Their advantages include:
-	\begin{itemize}
-		\item Simplicity.
-		\item Computational efficiency.
-		\item Domain and density adaptivity.
-		\item Practical success.
-	\end{itemize}
-	\item \textbf{Our contributions, short version.} For these reasons, there is a  substantial body of work assessing the statistical properties of graph Laplacian based methods.
-\textcolor{red}{(TODO): Include citations.} In this paper, we adopt a different perspective, and analyze~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} through the minimax lens.
-We show that for a variety of function classes $\mc{F}$, distributions $P$, and choices of domain $\mc{X}$ for which minimax rates are known, both~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are minimax optimal estimators (with loss measured in a suitable norm) and both~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} are minimax optimal test statistics.
-This lends additional theoretical support to the case for using graph Laplacian methods.
-\end{itemize}
-
-\section{Problem Setup and Background}
-\label{sec:problem_setup_and_background}
-
-Before we can state our main results, we must formally define minimaxity in both the estimation and testing problems.
-As we do so, we will remind the reader of the minimax optimal rates for some classical function classes $\mc{F}$ that we will henceforth consider.
-We will also briefly recall some (of the many) estimators and tests which achieve these rates; in particular we will focus on the \emph{truncated series} and \emph{smoothing spline} methods, which can be viewed as analogous to the graph Laplacian methods we consider.
-
-\begin{itemize}
-	\item \textbf{Minimax estimation.} For given integers $s > 0$ and $p > 0$, the Sobolev space $W^{s,p}(P)$ consists of all functions $f \in \Leb^p(P)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d)$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^p(P)$.
-The Sobolev $\{s,p\}$ norm is then
-	\begin{equation*}
-	\norm{f}_{W^{s,p}(P)}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^p \,dP
-	\end{equation*}
-	and for a given $M > 0$, the corresponding ball is $W^{s,p}(P, M) = \set{f: \norm{f}_{W^{s,p}(P)} \leq M}$.
-In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space; we adopt the usual convention of writing $H^s(P) = W^{s,p}(P)$ and $H^s(P,M) = W^{s,2}(P,M)$.
-We will confine our attention to this special case hereafter.
-Let us also consider the following assumption
-
-	\begin{enumerate}[(P1)]
-		\item 
-		\label{asmp:bounded_density}
-		The distribution $P$ is defined on domain $\Xset = [0,1]^d$, and admits a density $p$ with respect to the uniform (Lebesgue) measure on $[0,1]^d$.
-The density $p$ is bounded away from $0$ and $\infty$,
-		\begin{equation*}
-		0 < p_{\min} < p(x) < p_{\max} < \infty
-		\end{equation*}
-		for all $x \in [0,1]^d$.
-	\end{enumerate}
-	The assumption~\ref{asmp:bounded_density} is a standard but significant assumption; we will postpone a more significant discussion of it until \textcolor{red}{a later point}.
-Assuming~\ref{asmp:bounded_density}, the minimax estimation rate over the Sobolev ellipsoids $H^{s}(P,M)$ is
-	\begin{equation*}
-	\inf_{\wh{f}} \sup_{f \in H^s(P,M)} \Ebb\Bigl[\norm{\wh{f} - f}_{L^2(P)}^2\Bigr] \asymp n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d$,}
-	\end{equation*}
-	where the infimum is over all measurable maps $\wh{f}(X_1,Y_1,\ldots,X_n,Y_n)$
-	\item \textbf{Minimax hypothesis testing.}
-	\item \textbf{Our contributions, long version.}
-	\begin{itemize}
-		\item \textbf{Minimax optimal estimation.}
-		We show that with high probability,  $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2s/(2s + d)}$ whenever $f_0 \in C^s(\mc{X})$ for any $s \leq 1$, or $f_0 \in H^s(P)$ for $s = 1$.
-The same conclusions hold with respect to $\wt{f}_{\LS}$ when $d \leq 4$.
-		\item \textbf{Minimax optimal testing.}
-		We show that with high probability, a level-$\alpha$ test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in C^s(\mc{X})$ for any $s \leq 1$---or $f_0 \in H^s(P)$ for $s = 1$---additionally satisfies $\norm{f_0}_{\Leb^2(P)} \gtrsim \max\{n^{-4s/(4s + d)},n^{-1/4}\}$.
-\textcolor{red}{The same conclusions hold with respect to $\wh{f}_{LS}$.}
-		\item \textbf{Manifold adaptivity.}
-		If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$, each of the aforementioned rates hold with $d$ replaced by $m$.
-		\item \textbf{Taking advantage of higher-order smoothness conditions.} Under appropriate boundary conditions on the function classes $C^s$ and $H^s$ and additional smoothness conditions on $P$, when $s > 1$  each of the aforementioned conclusions hold for certain values of $d$.
-	\end{itemize}
-	\item \textbf{Organization.}
-	\item \textbf{Notation.}
-\end{itemize}
-
-\section{Minimax optimal graph Laplacian methods}
-
-Let us start by assuming $f_0 \in H^1(P,M)$.
-In this case, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and the resulting test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test}, achieve the standard minimax optimal estimation and hypothesis testing rates, respectively.
-
-\subsection{Minimax optimal estimation}
-
-Under appropriate conditions, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves minimax rates over all $f_0 \in H^1(P,M)$ with high probability.
-The conditions we impose will be on the neighborhood graph $G_{n,r}$.
-As the construction of the graph is in the user's control, these assumptions are therefore mild.
-\begin{enumerate}[label=(K\arabic*)]
-	\item 
-	\label{asmp:kernel_form}
-	The kernel function $K(z) = \1\Bigl\{\norm{z}_2/r \leq 1 \Bigr\}$.
-	\item 
-	\label{asmp:kernel_radius}
-	The radii $r = r(n)$ satisfies $r(n) \to 0$ as $n \to \infty$, and additionally
-	\begin{equation*}
-	n^{-1/d}\leq r(n) \leq M^{2/(2 + d)}n^{-1/(2 + d)}
-	\end{equation*}
-	for all $n$ sufficiently large.
-\end{enumerate}
-
-We use~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius} to ensure that the approximation error (bias) of $\wh{f}_{\LE}$ is of the correct order.
-On the other hand, bounding the estimation error of $\wh{f}_{\LE}$ -- both in expectation and in probability-- is straightforward.
-Together, bounds on these two sources of error imply Theorem~\ref{thm:laplacian_eigenmaps_estimation1}.
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_estimation1}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
-Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius}.
-Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
-	\begin{equation*}
-	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4 + d)} n^{-2/(2 + d)}
-	\end{equation*}
-	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
-\end{theorem}
-
-\begin{itemize}
-	\item Discuss the choice of radius $r$ imposed by~\ref{asmp:kernel_radius}.
-	\item Comment on the distinction between error rates as measured by the norm in $L^2(P_n)$ vs. $L^2(P)$. 
-\end{itemize}
-
-Theorem~\ref{thm:laplacian_smoothing_estimation1} is analogous to Theorem~\ref{thm:laplacian_eigenmaps_estimation1}, dealing with the Laplacian smoothing estimator $\wt{f}_{\LS}$.
-The key difference is that it holds only when the dimension $d < 4$.
-\begin{theorem}
-	\label{thm:laplacian_smoothing_estimation1}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
-Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius}.
-Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
-	\begin{equation*}
-	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4 + d)} n^{-2/(2 + d)}
-	\end{equation*}
-	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
-\end{theorem}
-
-\begin{itemize}
-	\item A restriction on the maximum value of dimension $d$ such as $d \leq 4$ is unsurprising.
-In a related context (estimation over a $d$-dimensional lattice graph), \cite{sadhanala2016} require the same restriction.
-The underlying phenomenon is similar to that elucidated by the seminal papers~\cite{birge1993} and~\cite{birge1998}: when the dimension $d$ is too large, the entropy function is irregular and (penalized) least squares approaches are no longer minimax optimal.
-\end{itemize}
-
-\subsection{Minimax optimal hypothesis testing}
-
-\section{Higher-order smoothness classes.}
-\label{sec:higher_order_smoothness_classes}
-Suppose we are willing to assume $f_0$ possesses more than one smooth derivative.
-We can show that in this situation, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} obtain the faster (known) minimax rates in the nonparametric estimation and testing problems.
-However, these conclusions hold only for certain values of the dimension $d$.
-
-\subsection{Minimax optimal estimation}
-Under  conditions, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves minimax rates over all $f_0 \in H_0^s(P,M)$ with high probability.
-These conditions will take two forms.
-The first pair of conditions will be on the neighborhood graph $G_{n,r}$.
-These assumptions are mild, as the construction of the graph is in the user's control.
-\begin{enumerate}[label=(K\arabic*)]
-	\item 
-	\label{asmp:kernel_form}
-	The kernel function $K(z) = \1\{\norm{z}_2/r \leq 1 \}$.
-	\item 
-	\label{asmp:kernel_radius}
-	The radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
-	\begin{equation*}
-	n^{-1/(2(s - 1) + d)}\leq r(n) \leq M^{2/(2s + d)}n^{-1/(2s + d)}
-	\end{equation*}
-	for all $n$ sufficiently large.
-\end{enumerate}
-
-We will also need to make an additional assumption on $P$ (additional to~\ref{asmp:bounded_density}, that is).
-The assumption we will make is stronger than~\ref{asmp:bounded_density}, but we need it to make $\Lap_{{n,r}}$---or more precisely, the iterated Laplacian operator $\Lap_{{n,r}}^s$---behave \textcolor{red}{like an appropriate differential operator}.
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{1}
-	\item 
-	\label{asmp:smooth_density}The distribution $P$ satisfies~\ref{asmp:bounded_density}. Additionally, the density $p \in C^{\beta}([0,1]^d;p_{\max})$ for some $\beta \geq 0$. 
-\end{enumerate}
-Together these assumptions are used to ensure that the approximation error (bias) of $\wh{f}_{\LE}$ is of the correct order.
-Bounds on the estimation error of $\wh{f}_{\LE}$ -- both in expectation and in probability-- are straightforward.
-Together, bounds on these two sources of error imply Theorem~\ref{thm:laplacian_eigenmaps_estimation}.
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_estimation}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
-Suppose that $f_0 \in H_0^s(P,M)$, that $P$ satisfies~\ref{asmp:smooth_density} with $\beta = s - 1$, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius}.
-Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as in~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2s + d)}n^{d/(2s + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
-	\begin{equation*}
-	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4s + d)} n^{-2s/(2s + d)}
-	\end{equation*}
-	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
-\end{theorem}
-
-\begin{enumerate}
-	\item \textcolor{red}{(TODO)}: Comment on the difference between this and minimax optimality. Cite others \textcolor{red}{(Wainwright, Liu)} who do the same thing.
-	\item \textcolor{red}{(TODO)}: Comment on why we insist that $f_0$ satisfy a boundary condition.
-	\item \textcolor{red}{(TODO)}: Comment a bit on the proof strategy; i.e., how we don't convergence properties of the eigenvectors $v_k$ towards eigenfunctions of the weighted Laplace-Beltrami operator, why we think that's advantageous, and how it might be limiting.
-\end{enumerate}
-
-\section{Fixed graph results}
-
-\section{Simulations}
-\begin{itemize}
-	\item \textbf{Estimation MSE 1.} Plot $L^2(P_n)$ error as a function of $n$ and choice $f_0^{(n)}$ when $P$ is (close to) uniform, showing that Laplacian methods are competitive with their non-graph-based brethren.
-	\item \textbf{Estimation fitted values.} Show fits for two regression problems, one where the distribution $P$ is assumed to be close to uniform, one where $P$ is dramatically non-uniform.
-	\item \textbf{Estimation MSE 2.} Plot $L^2(P_n)$ error as a function of the distribution $P$, as we vary $P$ from being uniform to being very non-uniform.
-
-	\item \textbf{Testing power.} 
-\end{itemize}
-
-\begin{figure}
-	\centering
-	\includegraphics[width = .48\textwidth]{plots/graph_regression/mean_squared_errors/sobolev_MSE_by_density}
-	\caption{}
-	\label{fig:piecewise_cosine}
-\end{figure}
-
-\clearpage
-
-\appendix
-
-Many of our results follow the same general two-part proof strategy.
-First, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to functionals defined on $G$.
-We then analyze the behavior of these functionals on the particular neighborhood graph $G_{n,r}$, giving high probability bounds which hold given our assumptions on the distribution $P$.
-
-\section{Fixed graph error bounds}
-
-\subsection{Laplacian eigenmaps on a fixed graph}
-Suppose we observe $G = ([n],E)$ a fixed graph, and responses
-\begin{equation}
-\label{eqn:fixed_graph_regression_model}
-Y_i = f_{0,i} + \varepsilon_i
-\end{equation}
-where $f_0 \in \Reals^n$ and $\varepsilon_i$ are independent $N(0,1)$ noise. The Laplacian eigenmaps estimator of $f_0$ on $G$ is then
-\begin{equation}
-\label{eqn:laplacian_eigenmaps_G}
-\wh{f}_{\LE}(G) := \sum_{k = 1}^{\kappa} \biggl(\frac{1}{n}\sum_{i = 1}^{n} Y_i v_{k,i}(G) \biggr) v_k(G),
-\end{equation}
-where we note that \eqref{eqn:laplacian_eigenmaps} specializes~\eqref{eqn:laplacian_eigenmaps_G} to the case where $G = G_{n,r}$.
-
-In Lemma~\ref{lem:fixed_graph_estimation}, we upper bound the mean squared error of $\wh{f}_{\LE}(G)$.
-Our bound will be stated as a function of $f_0^T \Lap_G^s f_0$--a measure of the smoothness the signal $f_0$ displays over the graph $G$--as well as the $\kappa$th eigenvalue $\lambda_{\kappa}(G)$.
-
-\begin{lemma}
-	\label{lem:fixed_graph_estimation}
-	Suppose we observe a graph $G$ and responses $Y$ according to the model~\eqref{eqn:fixed_graph_regression_model}. Then for any $1 \leq \kappa \leq n$, 
-	\begin{equation}
-	\label{eqn:fixed_graph_estimation}
-	\Ebb\Bigl[\norm{\wh{f}_{\LE}(G) - f_0}_n^2\Bigr] \leq 2\biggl(\frac{f_0^T \Lap_{G}^s f_0}{n \bigl(\lambda_{\kappa}(G)\bigr)^s} + \frac{\kappa}{n}\biggr)
-	\end{equation}
-	and
-	\begin{equation}
-	\label{eqn:fixed_graph_estimation_1}
-	\norm{\wh{f}_{\LE}(G) - f_0}_n^2 \leq 2\biggl(\frac{f_0^T \Lap_{G}^s f_0}{n \bigl(\lambda_{\kappa}(G)\bigr)^s} + \frac{4\kappa}{n}\biggr)
-	\end{equation}
-	with probability at least $1 - \exp(-\kappa)$.
-\end{lemma}
-The two terms on the right hand side of~\eqref{eqn:fixed_graph_estimation} and~\eqref{eqn:fixed_graph_estimation_1} represent the squared bias and variance of the estimator $\wh{f}_{\LE}(G)$.
-Note that the squared bias--- and therefore the overall mean squared error---depends on the quadratic form $f_0^T \Lap_{G}^s f_0$ and the eigenvalue $\lambda_{\kappa}(G)$, whereas the variance depends instead on the dimension $\kappa$ of the subspace $V_{\kappa} = \textrm{span}\{v_1,\ldots,v_{\kappa}\}$ to which $\wh{f}_{\LE}(G)$ belongs.
-In this respect, the error bound on the Laplacian eigenmaps estimator resembles the error bound on the \textcolor{red}{orthogonal series} estimator (see e.g.
-\textcolor{red}{Tsybakov}).
-
-To prove Theorems~\ref{thm:laplacian_eigenmaps_estimation},\textcolor{red}{...}, we would like to apply Lemmas~\ref{lem:fixed_graph_estimation},\textcolor{red}{...} in the random design context, i.e.
-to the random geometric graph $G_{n,r}$.
-To do so, we must establish the following (high probability) bounds: (i) an upper bound on the graph Sobolev seminorm of the regression function $f_0 \Lap_{n,r}^s f_0$, (ii) and lower boud on the Laplacian eigenvalue $\lambda_{\kappa}(G_{n,r})$ as a function of $\kappa$, and (iii) a lower bound on the empirical norm $\norm{f}_{n}$ as a function of $\norm{f}_{L^2(P)}$.
-
-\section{Bounds on the graph Sobolev semi-norm}
-
-\section{Bounds on graph eigenvalues}
-
-\section{Proofs of Theorems}
-\label{sec:proofs_theorems}
-
-\subsection{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation}}
-\label{subsec:laplacian_eigenmaps_estimation_pf}
-
-\paragraph{\eqref{eqn:fixed_graph_estimation_sobolev_rate_2} holds with high probability.}
-
-We first bound $f_0^T \Lap_{n,r}^s f_0$ in expectation.
-\begin{lemma}
-	\label{lem:roughness_functional_expectation_sobolev}
-	Let $\Xset$ be a Lipschitz domain and let $s \geq 1$ be an integer.
-Suppose that \textcolor{red}{$f \in H_0^{s}(P,M)$}, and that $P$ admits a density $p \in C^{s-1}(\Xset,p_{\max})$ for some constant $p_{\max} < \infty$.
-Let $\{r(n)\}$ be a sequence of connectivity radii such that $n^{-1/(2(s - 1) + d)} \leq r(n)$ for all $n$ sufficiently large, and $r(n) \to 0$ as $n \to \infty$.
-Then \textcolor{red}{for any $2$nd-order kernel $K$}, there exists a constant $c$ depending only on $p_{\max}, K_{\max}$, $s$, $d$ and $\Xset$ such that the resulting neighborhood graph Laplacian satisfies
-	\begin{equation}
-	\label{eqn:roughness_functional_expectation_sobolev}
-	\Ebb\bigl[f^T L_{n,r}^s f\bigr] \leq c \cdot n \cdot (nr^{d+2})^s \cdot M^2
-	\end{equation}
-	for all $n$ sufficiently large.
-\end{lemma}
-
-Since $\Lap_{n,r}$ is a positive semi-definite matrix, $f_0^T \Lap_{{n,r}}^s f_0$ must be non-negative. We may therefore apply Markov's inequality and conclude
-\begin{equation}
-\label{eqn:roughness_functional_probability_sobolev}
-f_0^T \Lap_{{n,r}}^s f_0 \leq  \frac{1}{\delta} \cdot c \cdot n \cdot (nr^{d+2})^s \cdot M^2
-\end{equation}
-with probability at least $1 - \delta$.
-
-\paragraph{\eqref{eqn:fixed_graph_estimation_sobolev_rate_1} holds with high probability.}
-
-Lemma~\ref{lem:neighborhood_eigenvalue} establishes a scaling rate which holds for all eigenvalues $\lambda_{k}(G_{n,r})$, assuming appropriate regularity conditions on $P$.
-\begin{lemma}
-	\label{lem:neighborhood_eigenvalue}
-	Suppose the distribution $P$ satisfies~\ref{asmp:bounded_density}.
-Then there exists constants $c$ and $C$ which depend only on $p_{\min}, p_{\max}$ and $d$, such that with probability at least $1 - C r^{-d} \exp\bigl\{- c n r^d\bigr\}$, for all $2 \leq k \leq n$
-	\begin{equation}
-	\label{eqn:neighborhood_eigenvalue}
-	\lambda_k(G_{n,r}) \geq c \cdot \min\Bigl\{nr^{d + 2}k^{2/d}, nr^d\Bigr\}.
-	\end{equation}	
-\end{lemma}
-The first term on the right hand side of~\eqref{eqn:neighborhood_eigenvalue} is the scaling of the eigenvalues we desire.
-The second term reflects the fact that the maximum eigenvalue of the Laplacian can never be more than twice the maximum degree of the graph.
-It is a bit of simple algebra to derive the conditions under which the first term is the smaller of the two when $k = \kappa = M^{2/(2s + d)} n^{d/(2s + d)}$.
-\begin{corollary}
-	\label{cor:neighborhood_eigenvalue}
-	Let $\kappa = M^{2/(2s + d)} n^{d/(2s + d)}$. For any $r \leq M^{-2/(d(2s +d))} n^{-1/(2s + d)}$, 
-	\begin{equation}
-	\label{eqn:neighborhood_eigenvalue_1}
-	\lambda_{\kappa}(G_{n,r}) \geq c \cdot nr^{d+2}{\kappa}^{2/d}
-	\end{equation}
-	with probability at least $1 - C r^{-d} \exp\{- c n r^d\}$. The constants $c$ and $C$ are the same as in Lemma~\ref{lem:neighborhood_eigenvalue}.
-\end{corollary}
-
-\paragraph{Putting the pieces together.}
-
-Let $n$ be sufficiently large so that $n^{-1/(2(s - 1)+ d)} \leq r(n) \leq M^{-2/(d(2s + d))} n^{-1/(2s + d)}$.
-Then by~\eqref{eqn:roughness_functional_probability_sobolev} and~\eqref{eqn:neighborhood_eigenvalue_1}, there exists a set $\mathcal{E}(f_0) \subseteq \Xset^n$ of probability at least $1 - \delta - Cr^{-d}\exp\{-cnr^d\}$ such that whenever ${\bf X} \in \mathcal{E}(f_0)$,
-\begin{equation*}
-f_0^T \Lap_{n,r}^s f_0 \leq C \cdot n \cdot C_n^s \cdot M^2
-\end{equation*}
-and
-\begin{equation*}
-\lambda_{\kappa}(G_{n,r}) \geq c \cdot nr^{d+2}\kappa^{2/d}
-\end{equation*}
-Theorem~\ref{thm:laplacian_eigenmaps_estimation} then follows by Corollary~\ref{cor:fixed_graph_estimation_sobolev_rate}.
-
-\bibliographystyle{plainnat}
-\bibliography{../graph_testing_bibliography}
-
 \end{document}

### 2020-08-14 notes/weekly_notes/8.12.20.NotesForWeek.tex
8.12.20.NotesForWeek. Added fixed graph testing results for Laplacian smoothing test.

@@
 \newcommand{\Reals}{\mathbb{R}}
-\newcommand{\Identity}{\mathbb{I}}
 \newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
@@
 \newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}
-\newcommand{\Lap}{{\bf L}}
+\newcommand{\Lap}{L}
 \newcommand{\NLap}{{\bf N}}
 \newcommand{\PLap}{{\bf P}}
+\newcommand{\Id}{I}
 
@@
 
+\section{Fixed graph regression}
+
+Suppose we observe a graph $G = \bigl([n],W\bigr)$, and responses
+\begin{equation}
+\label{eqn:fixed_graph_regression_model}
+Y_i = \theta_{0,i} + \varepsilon_i
+\end{equation}
+with signal vector $\theta_0 = (\theta_{0,1},\ldots,\theta_{0,n}) \in \Reals^n$, and noise vector $\varepsilon = (\varepsilon_1,\ldots,\varepsilon_n) \sim N(0,\Id_{n \times n})$.
+Letting $L$ denote the Laplacian matrix of graph $G$, the \emph{Laplacian smoothing} estimator $\wt{\theta}(G) \in \Reals^n$ is given by
+\begin{equation}
+\label{eqn:ls_G}
+\wt{\theta}(G) := \argmin_{\theta \in \Reals^n} \biggl\{ \sum_{i = 1}^{n}(Y_i - \theta_i)^2 + \rho \cdot \theta^T \Lap_G^{s}  \theta \biggr\} = (\rho \Lap_G^s + I)^{-1}Y.
+\end{equation}
+Now suppose we wish to test 
+\begin{equation}
+\mathbf{H}_0: \theta_0 = 0 ~~\textrm{vs.}~~ \mathbf{H}_a: \theta_0 \neq 0
+\end{equation}
+
+ A natural candidate is the quadratic form
+\begin{equation}
+\label{eqn:ls_ts_G}
+\wt{T}(G) := \frac{1}{n}\sum_{i = 1}^{n} \bigl(\wt{\theta}_i(G)\bigr)^2
+\end{equation}
+In Lemma~\ref{lem:ls_fixed_graph_testing}, we provide concentration bounds on the test statistic $\wt{T}(G)$, under both the null and alternative hypotheses.
+These statements are a function of the spectral decomposition $\Lap_G = \sum_{k = 1}^{n} \lambda_k(G) \cdot v_k(G) v_k(G)^T$.
+\begin{lemma}
+	\label{lem:ls_fixed_graph_testing}
+	Define the threshold $\wt{t}_b$ to be
+	\begin{equation*}
+	\wt{t}_b := \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k^s + 1\bigr)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k^s + 1\bigr)^4}}
+	\end{equation*}
+	Then,
+	\begin{itemize}
+		\item \textbf{Type I error.}
+		\begin{equation}
+		\label{eqn:ls_fixed_graph_testing_typeI}
+		\Pbb_0\Bigl(\wt{T}(G) > \wt{t}_b\Bigr) \leq \frac{1}{b^2}
+		\end{equation}
+		\item \textbf{Type II error.} If
+		\begin{equation}
+		\label{eqn:ls_fixed_graph_testing_critical_radius}
+		\frac{1}{n}\norm{\theta_0}_2^2 \geq \frac{2 \rho}{n} \bigl(\theta_0^T \Lap^s \theta_0\bigr) + \frac{4b}{n} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k^s + 1)^4} \Biggr)^{1/2}
+		\end{equation}
+		then
+		\begin{equation}
+		\label{eqn:ls_fixed_graph_testing_typeII}
+		\Pbb_{\theta_0}\Bigl(\wt{T}(G) \leq \wt{t}_b\Bigr) \leq \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k^s + 1)^4} \Biggr)^{-1/2}
+		\end{equation}
+	\end{itemize}
+\end{lemma}
+
+\subsection{Notation}
+\begin{itemize}
+	\item When the graph $G$ is obvious from context, we may drop the notational dependence on $G$ and simply write the Laplacian of $G$ as $\Lap$, and the $k$th eigenvalue of the Laplacian as $\lambda_k$.
+\end{itemize}
+
+\section{Proofs}
+
+\subsection{Proof of Lemma~\ref{lem:ls_fixed_graph_testing}}
+
+Let $\wt{S} := (\rho \Lap^s + \Id)^{-1}$.
+The matrix $\wt{S} \in \Reals^{n \times n}$ is symmetric and positive semidefinite, and $\wt{T}(G) = \frac{1}{n}Y^T \wt{S}^2 Y$.
+The desired result thus follows from Lemma~\ref{lem:linear_smoother_fixed_graph_testing}.
+To see that the conditions of this Lemma are satisfied, we first note that since
+\begin{equation*}
+\lambda_k(S) = \frac{1}{(\rho\lambda_k^s + 1)}
+\end{equation*}
+and $\rho, \lambda_k > 0$, it is evident that $\lambda_{\max}(\wt{S}) \leq 1$.  Then, by assumption~\eqref{eqn:ls_fixed_graph_testing_critical_radius}
+\begin{equation*}
+\theta_0^{T} \wt{S}^2 \theta_0 = \norm{\theta_0}_2^2 - \theta_0^T(I - \wt{S}^2)\theta_0 \geq 2 \rho (\theta_0^T \Lap^s \theta_0) + \theta_0^T(I - \wt{S}^2)\theta_0 + 4b \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k^s + 1)^4} \Biggr)^{-1/2},
+\end{equation*}
+and along with the following calculations,
+\begin{equation*}
+\begin{aligned}
+\theta_0^T \Bigl(\Id - \wt{S}^2\Bigr) \theta_0  & \overset{(i)}{=} \theta_0^T L^{s/2} L^{-s/2}\Bigl(\Id - \wt{S}^2\Bigr) L^{-s/2} L^{s/2} \theta_0 \\ 
+& \leq \theta_0^T L^{s} \theta_0 \cdot  \lambda_{\max}\biggl(L^{-s/2}\Bigl(\Id - \wt{S}^2\Bigr) L^{-s/2}\biggr) \\ 
+& \overset{(ii)}{=}  \theta_0^T L^{s} \theta_0 \cdot \max_{k} \biggl\{ \frac{1}{\lambda_k^s} \Bigl(1 - \frac{1}{(\rho \lambda_k^s + 1)^2}\Bigr) \biggr\} \\
+& \overset{(iii)}{\leq} \theta_0^T L^{s} \theta_0 \cdot 2\rho,
+\end{aligned}
+\end{equation*}
+we have that
+\begin{equation*}
+\theta_0^{T} \wt{S}^2 \theta_0 \geq 2b \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k^s + 1)^4} \biggr)^{-1/2}.
+\end{equation*} 
+In other words condition~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} in Lemma~\ref{lem:linear_smoother_fixed_graph_testing} is met, and applying that Lemma completes the proof.
+
+(In the previous derivation: in $(i)$ when we write $\Lap^{-s/2}$ we are referring to the pseudoinverse of $\Lap$, since the Laplacian always has at least one eigenvalue equal to $0$; in $(ii)$ the maximum is over all indices $k$ such that the eigenvalue $\lambda_k$ is strictly positive; and $(iii)$ follows from the basic algebraic identity $1 - 1/(1 + \rho x)^2 \leq 2 \rho x$ for any $x, \rho > 0$.
+
+\section{Technical Lemmas}
+
+\subsection{Type I and Type II error of quadratic forms.}
+Let $S \in \Reals^{n \times n}$ be a square symmetric matrix. The quadratic form
+\begin{equation}
+T = Y^T S^2 Y
+\end{equation}
+can be used as a test statistic to distinguish $\mathbf{H}_0$ from $\mathbf{H}_a$.
+In Lemma~\ref{lem:linear_smoother_fixed_graph_testing}, we establish conditions under which a test based on $T$ has small Type I and Type II error.
+\begin{lemma}
+	\label{lem:linear_smoother_fixed_graph_testing}
+	Define the threshold $t_b$ to be 
+	\begin{equation}
+	t_{b} := \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^2 + 2b \sqrt{\sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^4}
+	\end{equation}
+	Suppose the eigenvalues $0 \leq \lambda_{\min}(S) \leq \lambda_{\max}(S) \leq 1$. Then,
+	\begin{itemize}
+		\item \textbf{Type I error.}
+		\begin{equation}
+		\label{eqn:linear_smoother_fixed_graph_testing_typeI}
+		\Pbb_0\bigl(T > t_b\bigr) \leq \frac{1}{b^2}
+		\end{equation}
+		\item \textbf{Type II error.} Assuming that
+		\begin{equation}
+		\label{eqn:linear_smoother_fixed_graph_testing_critical_radius}
+		\theta_0^T S^2 \theta_0 \geq 4b \sqrt{\sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^4}
+		\end{equation}
+		then
+		\begin{equation}
+		\label{eqn:linear_smoother_fixed_graph_testing_typeII}
+		\Pbb_{\theta_0}\bigl(T \leq t_b\bigr) \leq \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n}\bigl(\lambda_k(S)\bigr)^4\Biggr)^{-1/2}
+		\end{equation}
+	\end{itemize}
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:linear_smoother_fixed_graph_testing}]
+	We compute the mean and variance of $T$ as a function of $\theta_0$, then apply Chebyshev's inequality.
+	\paragraph{Mean.}
+	Writing $Y = \theta_0 + \varepsilon$, we make use of the eigendecomposition $S = \sum_{k = 1}^{n} \lambda_k(S) \cdot v_k(S) v_k(S)^T$---where in this case we fix the eigenvectors $v_1(S),\ldots,v_n(S)$ to be unit-norm---and obtain
+	\begin{equation}
+	\label{pf:linear_smoother_fixed_graph_testing1}
+	\begin{aligned}
+	T & = \theta_0^T S^2 \theta_0 + 2 \theta_0^T S^2 \varepsilon + \varepsilon^T S^2 \varepsilon \\
+	& = \theta_0^T S^2 \theta_0 + 2 \theta_0^T S^2 \varepsilon + \sum_{k = 1}^{n}  \bigl(\lambda_k(S)\bigr)^2 (\varepsilon^T v_k(S))^2 \\
+	& = \theta_0^T S^2 \theta_0 + 2 \theta_0^T S^2 \varepsilon + \sum_{k = 1}^{n}  \bigl(\lambda_k(S)\bigr)^2 Z_k^2
+	\end{aligned}
+	\end{equation}
+	where in the last line $Z_k = (\varepsilon^T v_k(S))$, and $Z = (Z_1,\ldots,Z_k) \sim N(0,\Id)$ follows from the rotational invariance of the Gaussian distribution. Thus
+	\begin{equation}
+	\label{pf:linear_smoother_fixed_graph_testing_mean}
+	\Ebb_{\theta_0}[T] = \theta_0^T S^2 \theta_0 + \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^2.
+	\end{equation}
+	\paragraph{Variance.}  
+	Starting from~\eqref{pf:linear_smoother_fixed_graph_testing1} and recalling the basic fact $\Var(Z_k^2) = 2$, we derive
+	\begin{equation}
+	\label{pf:linear_smoother_fixed_graph_testing_var}
+	\Var_{\theta_0}[T] \leq 8 \theta_0^T S^4 \theta_0 + 4 \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^4 \leq 8 \theta_0^T S^2 \theta_0 + 4 \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^4
+	\end{equation}
+	where the second inequality follows since by assumption $\lambda_{\max}(S) \leq 1$.
+
+	\paragraph{Bounding Type I and Type II error.}
+	The bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeI} follows directly from Chebyshev's inequality, along with our above calculations on the mean and variance of $T$.
+
+	The bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeII} also follows from Chebyshev's inequality, as can be seen by the following manipulations,
+	\begin{equation*}
+	\begin{aligned}
+	\Pbb_{\theta_0}\bigl(T \leq t_b\bigr) & = \Pbb_{\theta_0}\bigl(T - \Ebb_{\theta_0}[T] \leq t_b - \Ebb_{\theta_0}[T]\bigr) \\
+	& \overset{(i)}{\leq} \Pbb_{\theta_0}\bigl(\abs{T - \Ebb_{\theta_0}[T]} \geq \abs{t_b - \Ebb_{\theta_0}[T]}\bigr) \\ 
+	& \overset{(ii)}{\leq} 4 \frac{\Var_{\theta_0}[T]}{(\theta_0^T S^2 \theta_0)^2} \\
+	& \overset{(iii)}{\leq} \frac{32}{\theta_0^T S^2 \theta_0} + \frac{4}{b^2} \\
+	& \overset{(iv)}{\leq} \frac{8}{b} \Biggl(\sum_{k = 1}^{n}\bigl(\lambda_k(S)\bigr)^4\Biggr)^{-1/2} + \frac{4}{b^2}
+	\end{aligned}
+	\end{equation*}
+	In the previous expression, $(i)$ and $(ii)$ follow since assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and equation~\eqref{pf:linear_smoother_fixed_graph_testing_mean} together imply $\Ebb_{\theta_0}(T) - \frac{1}{2}\theta_0^T S^2\theta_0 \geq t_b$, $(iii)$ follows from assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and the inequality~\eqref{pf:linear_smoother_fixed_graph_testing_var}, and $(iv)$ follows assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}.
+\end{proof}
+
 \end{document}

### 2020-08-19 notes/weekly_notes/3.12.20.NotesForWeek.tex
Added weekly notes, which include some derivations of testing results for the fixed graph case.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\makeatletter
\newcommand{\vast}{\bBigg@{4}}
\newcommand{\Vast}{\bBigg@{5}}
\makeatother

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{\mathcal{L}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LS}{\mathrm{LS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 3/6/20 - 3/12/20}
\author{Alden Green}
\date{\today}
\maketitle

Suppose we observe independent design points $X = x_1,\ldots,x_n \sim P$ i.i.d -- where we assume $P$ has density $p$ which is supported on $[0,1]^d$ -- and responses
\begin{equation}
\label{eqn:regression_random_design_known_variance}
y_i = f(x_i) + \varepsilon_i, \varepsilon_i \overset{i.i.d}\sim N(0,1).
\end{equation} 
Our goal is to test
\begin{equation*}
\mathbf{H}_0: f = 0,~~\textrm{vs.}~~\mathbf{H}_a: f \neq 0.
\end{equation*}
We will use as our test statistic the empirical norm of a Laplacian smoothing estimator. Let $G = (X,E)$ be a graph formed over the design points $X$, the Laplacian smoothing estimator $\wh{\theta}_{\LS}(G) \in \Reals^n$ is defined as
\begin{equation}
\label{eqn:laplacian_smoothing_estimator}
\wh{\theta}_{\LS}(G) = \argmin_{\theta \in \Reals^n} \sum_{i = 1}^{n} (y_i - \theta_i)^2 + \rho \theta^T L_G \theta 
\end{equation}
i.e. $\wh{\theta}_{\LS}(G) = (I + \rho L_G)^{-1} y$; here $\rho > 0$ is a tuning parameter controlling how much shrinkage the estimator performs. Then our Laplacian smoothing test statistic will simply be
\begin{equation}
\label{eqn:laplacian_smoothing_statistic}
T_{\LS}(G) := \norm{\wh{\theta}_{\LS}(G)}_n^2.
\end{equation}
with the corresponding test $\phi_{\LS}(G) := \1\{T_{\LS}(G) \geq \tau(b) \}$, where $b > 1$ is a user specified hyperparameter which controls the level of Type I and Type II error the user is willing to tolerate, and $\tau$ is a function of $b$ (and also, implicitly, of $G$) to be specified later.

Let $G_{n,r}$ be the random geometric graph of radius $r$, i.e $G_{n,r} = (X,E_{n,r})$ where $E_{n,r} \subseteq X \times X$ contains the edge $e(x_i,x_j) \in E_{n,r}$ if and only if $\1(\norm{x_i - x_j}_2 \leq r)$. When $f \in H^1(\mathcal{X};L)$ for $d = 1,2$ or $3$, and the density $p$ satisfies typical regularity conditions, the test $\phi_{\LS}(G_{n,r})$ achieves minimax optimal testing rates.
\begin{theorem}
	\label{thm:sobolev_testing_rate_laplacian_smoothing_order1}
	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}. Let $L > 0$ and $b \geq 1$ be fixed constants, and $d = 1,2$ or $3$. Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p(x)$ bounded away from zero and infinity, 
	\begin{equation*}
	0 < p_{\min} < p(x) < p_{\max} < \infty,~~ \textrm{for all $x \in \mathcal{X}$.}
	\end{equation*}
	and the test $\phi_{\LS}(G_{n,r})$ is performed with parameter choices
	\begin{align*}
	c \frac{(\log n)^{p_d}}{n^{1/d}} \leq r(n) \leq n^{-4/((4 + d)(2+d))}& ,~~ \rho = \lambda_{\kappa}^{-1},~~ \kappa = n^{2d/(4 + d)}\\
	\tau(b) = \frac{1}{n} \sum_{k = 1}^{n} \Bigl(\frac{1}{1 + \rho \lambda_k}\Bigr)^2 & + \frac{4b}{n}\sqrt{\sum_{k = 1}^{n} \Bigl(\frac{1}{1 + \rho \lambda_k}\Bigr)^4}
	\end{align*}
	for $c$ a constant that depends only on $\Xset, p_{\min}$ and $p_{\max}$. Then the following statements holds for every $n$ sufficiently large: there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
	\begin{equation}
	\label{eqn:sobolev_testing_rate_order1}
	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot L^2 \cdot n^{-4/(4 + d)}
	\end{equation}
	the worst-case risk is upper bounded
	\begin{equation}
	\label{eqn:sobolev_testing_rate_order1_1}
	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}(G_{n,r}); H^1(\mathcal{X};L)) \leq \frac{c_2}{b}.
	\end{equation}
\end{theorem}

\section{Fixed Graph Analysis}
As usual, our analysis will proceed by showing that for any function $f$, there exists a set $E_f$ satisfying $\Pbb(E_f) \geq 1 - \norm{S_{\rho}(\beta)}_n^2$, such that conditional on $X = x$ for any $x \in E_f$ our test has non-trivial power. This latter step amounts to analyzing the behavior of our test in the fixed graph setting. Formally, suppose we observe fixed design points $x_1,\ldots,x_n$, and random responses 
\begin{equation}
\label{eqn:fixed_design_model}
y_i = \beta_i + \epsilon_i,~~ \epsilon_i \overset{\textrm{i.i.d}}{\sim} N(0,1)
\end{equation}
In the following Lemma, we bound the Type I and Type II error of $\phi_{\LS}(G)$. For convenience, we denote $S_{\rho} = (I + \rho L)^{-1}$.

\begin{lemma}
	\label{lem:fixed_graph_laplacian_smoothing_test}
	Fix $\rho > 0$. Suppose we observe data according to model~\eqref{eqn:fixed_design_model}, and perform the test $\phi_{\LS}(G)$ with threshold
	\begin{equation*}
	\tau(b) = \frac{1}{n} \sum_{k = 1}^{n} \Bigl(\frac{1}{1 + \rho \lambda_k}\Bigr)^2 + \frac{b}{n}\sqrt{\sum_{k = 1}^{n} \Bigl(\frac{1}{1 + \rho \lambda_k}\Bigr)^4}.
	\end{equation*}
	\begin{enumerate}
		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\LS}(G)$ is upper bounded,
		\begin{equation*}
		\Ebb_{\beta_0}\Bigl[\phi_{\LS}(G)\Bigr] \leq \frac{1}{b^2}
		\end{equation*}
		\item \textbf{Type II error:} 
		For any $b \geq 1$ and $\beta$ such that
		\begin{equation}
		\label{eqn:fixed_graph_laplacian_smoothing_test}
		\norm{S_{\rho}(\beta)}_n^2 \geq \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \Bigl(\frac{1}{1 + \rho \lambda_k}\Bigr)^4}
		\end{equation}
		the Type II error of $\phi_{\LS}(G)$ is upper bounded,
		\begin{equation}
		\label{eqn:fixed_graph_laplacian_smoothing_test2}
		\Ebb_{\beta}\Bigl[1 - \phi_{\LS}(G)\Bigr] \leq \frac{8}{b\sqrt{n}}\biggl(\sum_{k = 1}^{n} \Bigl(\frac{1}{1 + \rho \lambda_k}\Bigr)^4\biggr)^{-1/2} + \frac{2}{b^2}.
		\end{equation}
		In particular, there exist universal constants $c_1$ and $c_2$ such that if
		\begin{equation}
		\label{eqn:fixed_graph_laplacian_smoothing_radius}
		\frac{c_1}{n}\sum_{i = 1}^{n} \beta_i^2 \geq c_2\frac{\rho\beta^T L \beta}{n} + \frac{4b}{n}\sqrt{\sum_{k = 1}^{n} \Bigl(\frac{1}{1 + \rho \lambda_k}\Bigr)^4},
		\end{equation}
		then~\eqref{eqn:fixed_graph_laplacian_smoothing_test} and thus \eqref{eqn:fixed_graph_laplacian_smoothing_test2} follow.
	\end{enumerate}
\end{lemma}

\section{Analysis}

\subsection{Fixed Graph Testing}
Decomposing $y = \beta + \varepsilon$, the Laplacian smoothing test statistic may be written as
\begin{equation*}
T_{\LS}(G) = \frac{1}{n}(\beta + \varepsilon)^T S_{\rho}^{2} (\beta + \varepsilon) = \frac{1}{n} \Bigl(\beta^T S_\rho^{2} \beta + 2 \beta^T S_{\rho}^{2} \varepsilon + \varepsilon^T S_{\rho}^{2} \varepsilon \Bigr)
\end{equation*}

Writing the spectral decomposition of the Laplacian $L = V \Lambda V^T$ -- where $V$ is orthonormal and $\Lambda$ diagonal--and invoking the rotational invariance of the Gaussian distribution, we conclude that
\begin{equation}
\label{eqn:fixed_graph_ls_pf3}
\varepsilon^T S_{\rho}^2 \varepsilon \overset{d}{=} \sum_{k = 1}^{n} \Bigl(\frac{1}{1 + \rho \lambda_k}\Bigr)^2 \varepsilon_k^2
\end{equation}
This equality (in distribution) will be useful for computing both the mean and variance of $T_{\LS}(G)$.

\paragraph{Mean of $T_{\LS}(G)$.}

Noting that $\varepsilon$ is mean-zero, we have
\begin{align}
\Ebb\bigl[T_{\LS}(G)\bigr] & = \frac{1}{n}\Bigl(\beta^T S_\rho^{2} \beta + \Ebb\bigl[\varepsilon^T S_{\rho}^{2} \varepsilon\bigr] \Bigr) \nonumber \\
& = \norm{S_{\rho}(\beta)}_n^2 + \sum_{k = 1}^{n} \frac{1}{(1 + \rho \lambda_k)^2} \label{eqn:fixed_graph_ls_pf2}
\end{align}

\paragraph{Variance of $T_{\LS}(G)$.}
Note that since $\beta^T S_{\rho}^2 \varepsilon$ is symmetric about zero, 
\begin{equation*}
\Cov\bigl[\beta^T S_{\rho}^2 \varepsilon, \varepsilon^T S_{\rho}^2 \varepsilon\bigr] = 0,
\end{equation*}
and therefore
\begin{align*}
\Var\bigl[T_{\LS}(G)\bigr] & = \frac{1}{n^2}\Bigl(4 \Var\bigl[\beta^T S_\rho^2 \varepsilon \bigr] + \Var\bigl[\varepsilon^T S_{\rho}^2 \varepsilon\bigr]\Bigr) \\
& = \frac{1}{n^2}\Bigl(4 \beta^T S_{\rho}^4 \beta + \Var\bigl[\varepsilon^T S_{\rho}^2 \varepsilon\bigr]\Bigr) \\
& \leq \frac{1}{n^2}\Bigl(4 \beta^T S_{\rho}^2 \beta + \Var\bigl[\varepsilon^T S_{\rho}^2 \varepsilon\bigr]\Bigr) \\
& = \frac{1}{n^2}\biggl(4 \beta^T  S_{\rho}^2 \beta + \sum_{k = 1}^{n} \Bigl(\frac{1}{1 + \rho \lambda_k}\Bigr)^4 \Var\bigl[\varepsilon_k^2\bigr]\biggr) \\
& = \frac{1}{n^2}\biggl(4 \beta^T S_{\rho}^2 \beta + 2\sum_{k = 1}^{n} \Bigl(\frac{1}{1 + \rho \lambda_k}\Bigr)^4 \biggr).
\end{align*}
where the inequality in previous display follows from $\lambda_{\min}(S_{\rho}) \geq 1$.

For convenience, we introduce the notation 
\begin{equation*}
t(b) := \frac{b}{n}\sqrt{\sum_{k = 1}^{n} \Bigl(\frac{1}{1 + \rho \lambda_k}\Bigr)^4}.
\end{equation*}
\paragraph{Type I error.}
Using Chebyshev's inequality, we obtain
\begin{align*}
\Pbb_{\beta = 0}\Bigl(T_{\LS}(G) \geq \tau(b)\Bigr) & = \Pbb_{\beta = 0}\Bigl(T_{\LS}(G) - \Ebb\bigl[T_{\LS}(G)] \geq t(b)\Bigr) \\
& \leq \Pbb_{\beta = 0}\Bigl(\abs{T_{\LS}(G) - \Ebb\bigl[T_{\LS}(G)]} \geq t(b)\Bigr) \\
& \leq \frac{\Var_{\beta = 0}\bigl[T_{\LS}(G)\bigr]}{\bigl[t(b)\bigr]^2} \\
& \leq \frac{2}{b^2}.
\end{align*}

\paragraph{Type II error.}
We note that~\eqref{eqn:fixed_graph_ls_pf2} along with assumption~\eqref{eqn:fixed_graph_laplacian_smoothing_test} implies that
\begin{equation*}
\mathbb{E}[T_{\LS}(G)] - \tau(b) = \norm{S_{\rho}(\beta)}_n^2 - t(b) \geq t(b).
\end{equation*}
Again applying Chebyshev's inequality, we find
\begin{align*}
\Pbb_{\beta}\Bigl(T_{\LS}(G) < \tau(b)\Bigr) & = \Pbb_{\beta}\Bigl(T_{\LS}(G) -\Ebb_{\beta}\bigl[T_{\LS}(G)\bigr] < t(b) - \norm{S_{\rho}(\beta)}_n^2 \Bigr) \\
& \leq \Pbb_{\beta}\Bigl(\abs{T_{\LS}(G) -\Ebb_{\beta}\bigl[T_{\LS}(G)\bigr]} > \frac{1}{2}\norm{S_{\rho}(\beta)}_n^2  \Bigr) \\
& \leq 4\frac{\Var_{\beta}(T_{\LS}(G))}{\norm{S_{\rho}(\beta)}_n^4} \\
& \leq \frac{16 \norm{S_{\rho}(\beta)}_n^2/n + 8\sum_{k = 1}^{n} \Bigl(\frac{1}{1 + \rho \lambda_k}\Bigr)^4/n^2}{\norm{S_{\rho}(\beta)}_n^4}
\end{align*}
Since $\norm{S_{\rho}(\beta)}_n^2 \geq 2t(b)$,
\begin{equation*}
\frac{1}{n\norm{S_{\rho}(\beta)}_n^2} \leq \frac{1}{2nt(b)},~~ \frac{1}{n^2\norm{S_{\rho}(\beta)}_n^4}\sum_{k = 1}^{n} \Bigl(\frac{1}{1 + \rho \lambda_k}\Bigr)^4 \leq \frac{1}{4b^2}.
\end{equation*}
and~\eqref{eqn:fixed_graph_laplacian_smoothing_test2} follows.

\paragraph{\eqref{eqn:fixed_graph_laplacian_smoothing_radius} implies \eqref{eqn:fixed_graph_laplacian_smoothing_test}.}

Note that $S_{\rho}$ lets constant signals pass through unfiltered, i.e. decomposing $\beta = a_1 \wt{1} + a_2 \beta_{\perp}$ where $\wt{1} = n^{-1/2}(1,\ldots,1) \in \Reals^n$, we have
\begin{equation}
\label{eqn:fixed_graph_ls_pf1}
\beta^T S_\rho^{2} \beta = a_1^2 + a_2^2 \beta_{\perp}^T S_{\rho}^2 \beta_{\perp}
\end{equation}
We use the following crude but sufficient lower bound on the quadratic form,
\begin{equation*}
\beta_{\perp}^T S_\rho^{2} \beta_{\perp} = (\beta_{\perp} + (S_{\rho} - I)\beta_{\perp})^T (\beta_{\perp} + (S_{\rho} - I)\beta_{\perp}) \geq \Bigl(1 - \frac{1}{\sqrt{2}}\Bigr)\beta_{\perp}^T \beta_{\perp} - (4\sqrt{2} - 1) \Bigl(\beta^T (S_{\rho} - I)^T (S_{\rho} - I)\beta\Bigr).
\end{equation*}
Then the derivations in~\textcolor{red}{(Sadhanala)} show
\begin{equation*}
\beta^T (S_{\rho} - I)^T (S_{\rho} - I)\beta \leq \frac{\rho}{4} \beta^T L \beta
\end{equation*}
and plugging back in to~\eqref{eqn:fixed_graph_ls_pf1}, we conclude that
\begin{equation*}
\beta^T S_{\rho}^2 \beta \geq c_1 \beta^T \beta - c_2 \rho \beta^T L \beta 
\end{equation*}
for $c_1 = 1 - 1/\sqrt{2}$ and $c_2 = \sqrt{2} - 1/4$, which suffices to prove the claim.

\end{document}

### 2020-08-19 notes/weekly_notes/3.19.20.NotesForWeek.tex
Added weekly notes, which include some derivations of testing results for the fixed graph case.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\makeatletter
\newcommand{\vast}{\bBigg@{4}}
\newcommand{\Vast}{\bBigg@{5}}
\makeatother

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{\mathcal{L}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LS}{\mathrm{LS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 3/12/20 - 3/19/20}
\author{Alden Green}
\date{\today}
\maketitle

\section{Introduction}

Suppose we observe data according to the following random design regression sampling model: we observe independent samples $x_1,\ldots,x_n \sim P$ and responses
\begin{equation}
\label{eqn:regression_random_design_known_variance}
y_i = f(x_i) + \varepsilon_i,~~\varepsilon_i \sim N(0,1)
\end{equation}
where $f \in L^2(P)$ is a regression function and $\varepsilon$ is independent standard Gaussian noise. We wish to test
\begin{equation*}
\mathbf{H}_0: f = 0,~~ \mathbf{H}_a: f \neq 0.
\end{equation*}
Our strategy will be to estimate $\norm{f}_{\Leb^2(P)}$ using a test statistic $T$ to be specified momentarily, and reject the null hypothesis (which we can equivalently state as $\norm{f}_{\Leb^2} = 0$) if $T$ is sufficiently large.

The truncated series estimator of $f$,
\begin{equation*}
\Pi_{\kappa,G}(y) = \sum_{k = 1}^{\kappa} \biggl(\sum_{i = 1}^{n} v_{k,i}(G) y_i\biggr) v_k(G), 
\end{equation*}
leads to the somewhat natural choice of test statistic $T = \norm{\Pi_{\kappa,G}(y)}_n^2$. (Here, $G$ is a neighborhood graph built on the data $X$.)
A set of standard calculations lead to the following upper bound on approximation error:
\begin{equation*}
\abs{\norm{f}_{n}^2 - \norm{\Pi_{\kappa,G}(y)}_n^2} \leq \frac{f^T L^s f}{n \lambda_{\kappa}(G)^s}
\end{equation*}
When the graph $G$ is formed appropriately, the Laplacian matrix $L$ can be viewed as an estimate of a density-weighted Laplace-Beltrami operator $\Delta_P$. Then $\lambda_{\kappa}(G)$ can be viewed as an estimate of the $(\kappa)$th eigenvalue of $\Delta_P$, and $f^T L^s f$ as an estimate of a Sobolev semi-norm built via the operator $\Delta_P$. When the density $p$ satisfies appropriate regularity conditions, the population level quantities these estimates approximate are roughly
\begin{equation*}
\langle \Delta_P^s f, f \rangle_{\Leb^2(P)} \sim |f|_{H^{s}(\mathcal{X})}, ~~ \lambda_{\kappa}(G)^2 \sim \kappa^{-2s/d}
\end{equation*}
and the resulting approximation error incurred by projecting $f$ onto $\kappa$ eigenvectors in $\Leb^2(P_n)$ is comparable to that incurred by projecting $f$ onto $\kappa$ elements of the Fourier basis in $\Leb^2$.

Of course, this not a formal proof, and indeed when $s > 1$ we will see that problems arise. Let us assume for the moment that $f \in C^2(\mathcal{X};R)$.
In this case the second-order graph Sobolev seminorm $f^T L_{n,r}^2f$ is a biased estimator of the second-order Sobolev seminorm $\int (f^{(2)}(x))^2 \,dP(x)$. This bias will be non-trivial when $r$ is sufficiently small. On the other hand $\lambda_{\kappa}(G)$ is also a biased estimator of $\lambda_{\kappa}(\Delta_P)$, and this bias will be non-trivial when $r$ and $\kappa$ are sufficiently large. In particular, when $d > 4$, taking $\kappa = n^{2d/(8 + d)}$ as dictated by the usual bias-variance tradeoff in nonparametric testing problems, there exists no choice of $r$ for which the biases of both the seminorm $f^T L_{n,r}^2f$ and the eigenvalue $\lambda_{\kappa}(G)$ are trivial. A test formed using the statistic $\norm{\Pi_{\kappa,G}(y)}_n^2$ is therefore, at best, only minimax optimal when $d \leq 4$.

For this reason we introduce a bias correction of our statistic.

\subsection{Bias-Corrected Test Statistic}

Formally, define the (unweighted) neighborhood graph $G_{n,r} = (X,E)$, where $X = \set{x_1,\ldots,x_n}$, and $(x_i,x_j) \in E \subset X \times X$ if and only if $\norm{x_i - x_j}_2 \leq r$. Let $L_{n,r} := L_{G_{n,r}}$ be the Laplacian matrix associated with $G_{n,r}$. Our test statistic will be defined as a simple function of the norm of the projection of $Y$ onto the eigenvectors of $L_{n,r}$,
\begin{equation*}
T_{\spec}(G_{n,r}) := \norm{\Pi_{\kappa,G_{n,r}}(Y)}_n^2 + \frac{y^T L_{n,r} y}{2 n\bigl(\lambda_{\kappa}(G_{n,r})\bigr)^2}
\end{equation*}
The second term on the right hand side is the bias-correction term. Our test is then
\begin{equation*}
\phi_{\spec}(G_{n,r}) := \1\{T_{\spec}(G_{n,r}) \geq \tau(b) \}
\end{equation*}
where $b$ is a user-specified model encoding tolerance for error, and $\tau$ is a function of $b$ to be specified later.

The bias corrected graph spectral projection test is minimax optimal over the compactly supported Holder ball $C_0^2(\Xset;L)$ whenever $d < 8$. 
\begin{theorem}
	\label{thm:holder_testing_rate_2}
	Suppose we observe samples $(x_i,y_i)_{i = 1}^{n}$ according to the model~\eqref{eqn:regression_random_design_known_variance}. Let $L > 0$ and $b \geq 1$ be fixed constants, and $d < 8$. Suppose that $P$ is an absolutely continuous probability measure over $\mathcal{X} = [0,1]^d$ with density function $p(x) \in C^1(\Xset;p_{\max})$ bounded away from zero and infinity, 
	\begin{equation*}
	0 < p_{\min} < p(x) < p_{\max} < \infty,~~ \textrm{for all $x \in \mathcal{X}$.}
	\end{equation*}
	and the test $\phi_{\spec}(G_{n,r})$ is performed with parameter choices
	\begin{equation*}
	r(n) = n^{-2/(8 + d)}, ~\kappa = n^{2d/(4 + d)}, ~\tau(b) = \frac{\kappa}{n} + \frac{1}{2n\lambda_{\kappa}^2}\sum_{k = 1}^{n}\lambda_k + \sqrt{6}b\sqrt{\frac{2\kappa}{n^2}}
	\end{equation*}
	Then the following statements holds for every $n$ sufficiently large: there exists constants $c_1,c_2$ which do not depend on $n,b$ or $R$ such that for every $\epsilon \geq 0$ satisfying
	\begin{equation}
	\label{eqn:sobolev_testing_rate_order1}
	\epsilon^2 \geq c_1^2 \cdot b^2 \cdot L^2 \cdot n^{-8/(8 + d)}
	\end{equation}
	the worst-case risk is upper bounded
	\begin{equation}
	\label{eqn:sobolev_testing_rate_order1_1}
	\mathcal{R}_{\epsilon}(\phi_{\mathrm{spec}}(G_{n,r}); C_0^2(\mathcal{X};R)) \leq \frac{c_2}{b}.
	\end{equation}
\end{theorem}

\section{Analysis}

\subsection{Fixed graph testing error}

To prove Theorem~\ref{thm:holder_testing_rate_2}, we show that there exists a high-probability set $E \subseteq \Xset^n$ such that conditional on $X \in E$, the test $\phi_{\spec}(G_{n,r})$ has small risk. Since $G_{n,r}$ is a function only of $X$ and not of $Y$, this amounts to reasoning about the behavior of the test $\phi_{\spec}$ over a fixed graph $G = (X,E)$, where we observe
\begin{equation}
\label{eqn:fixed_graph_regression_model}
y_i = \beta_i + \varepsilon_i,~~\varepsilon_i \sim \mathcal{N}(0,1)
\end{equation}
for some fixed $\beta \in \Reals^n$.

In Lemma~\ref{lem:fixed_graph_testing}, we upper bound the Type I and Type II error of the test $\phi_{\spec}(G)$. Our bound on the Type II error will be stated as a function of $\beta^T L^2 \beta$--a measure of the smoothness the signal $\beta$ displays over the graph $G$--as well as the $\kappa$th eigenvalue $\lambda_{\kappa}$. (For ease of reading, in the statement and proof of this Lemma, we drop the notational dependence of the Laplacian $L$ and eigenvalues $\lambda$ on the graph $G$.)

\begin{lemma}
	\label{lem:fixed_graph_testing}
	Let $1 \leq \kappa \leq n$ be an integer. Suppose we observe data according to the model~\eqref{eqn:fixed_graph_regression_model}, and we perform the test $\phi_{\spec}(G)$ with
	\begin{equation*}
	\tau(b) = \frac{\kappa}{n} + \frac{1}{2n\lambda_{\kappa}^2}\sum_{k = 1}^{n}\lambda_k + 2\sqrt{6}\cdot b\sqrt{\frac{\kappa}{n^2}}.
	\end{equation*} 
	Assume that
	\begin{equation}
	\label{eqn:graph_spectral_1}
	\max\Bigl\{\beta^T L^2 \beta, n \lambda_n^2\Bigr\} \leq \kappa \cdot \lambda_{\kappa}^4
	\end{equation}
	\begin{enumerate}
		\item \textbf{Type I error:} Under the null hypothesis $\beta = \beta_0 = 0$, the Type I error of $\phi_{\spec}(G)$ is upper bounded
		\begin{equation}
		\label{eqn:graph_spectral_type_I_error}
		\mathbb{E}_{\beta_0}(\phi_{\spec}) \leq \frac{1}{b^2}.
		\end{equation}
		\item \textbf{Type II error:} For any $b \geq 1$ and $\beta$ such that
		\begin{equation}
		\label{eqn:fixed_graph_testing_critical_radius_1}
		\frac{\beta^T L \beta}{2 n \lambda_{\kappa}^2} + \norm{\Pi_{\kappa,G}(\beta)}_n^2 \geq 2 \sqrt{6} \cdot b\sqrt{\frac{\kappa}{n^2}}
		\end{equation}
		the Type II error of $\phi_{\spec}(G)$ is upper bounded,
		\begin{equation}
		\label{eqn:graph_spectral_type_II_error}
		\mathbb{E}_{\beta}(1 - \phi_{\spec}) \leq \frac{1}{b^2} + \frac{16}{\sqrt{6} \cdot b\sqrt{\kappa}}.
		\end{equation}
	\end{enumerate}
    In particular if
    \begin{equation}
    \label{eqn:fixed_graph_testing_critical_radius}
    \frac{1}{n} \sum_{i = 1}^{n} \beta_i^2 \geq 2\sqrt{6} \cdot b\sqrt{\frac{\kappa}{n^2}} + \frac{U_2(\beta;G)}{4 n\lambda_{\kappa}^2}
    \end{equation}
    where
    \begin{equation*}
    U_2(\beta;G) := \sum_{i \in (n)^3} (\beta_{i_1} - \beta_{i_2}) (\beta_{i_1} - \beta_{i_3}) A_{i_1i_2} A_{i_1i_3}
    \end{equation*}
    then~\eqref{eqn:fixed_graph_testing_critical_radius_1} and thus~\eqref{eqn:graph_spectral_type_II_error} follow.
\end{lemma}

\subsection{Variance of bias-correction term}

We pay a price in increased variance for introducing the bias-correction term $y^T L y/\bigl(2 n \lambda_{\kappa}^2\bigr)$; the condition~\eqref{eqn:graph_spectral_1} is sufficient to guarantee this increase in variance is no greater than the variance $\Var(||\Pi_{\kappa,G}(Y)||_n^2)$ of the uncorrected statistic. In this section, we show that under the conditions of Theorem~\ref{thm:holder_testing_rate_2}, the condition~\eqref{eqn:graph_spectral_1} holds with high probability.

\begin{lemma}
	\label{lem:bias_correction_variance}
	Fix $b \geq 1$. Suppose $f \in C_0^2(\Xset;L)$, and let $r = n^{-2/(8 + d)}$ and $\kappa = n^{2d/(8 + d)}$. Then there exist constants $c$ and $C$ such that for all $n$ sufficiently large and all $d < 8$,
	\begin{equation}
	\label{eqn:bias_correction_variance}
	\max\Big\{f^T L_{n,r}^2 f,~ n \bigl(\lambda_n(G_{n,r})\bigr)^2 \Bigr\} \leq \kappa \cdot \bigl(\lambda_{\kappa}(G_{n,r})\bigr)^4
	\end{equation}
	with probability at least $1 - 1/b - 1/(nr^4) - C(n + r^{-d}) \exp\{-c nr^d\}$.
\end{lemma}

\subsection{Approximation error}

The following Lemma is the main result of this section.
\begin{lemma}
	\label{lem:approximation_error}
	Set $r = n^{-2/(8 + d)}$ and $\kappa = n^{2d/(8 + d)}$. Then if $f \in C_0^2(\mathcal{X};L)$, $p \in C^1(\mathcal{X};p_{\max})$, and additionally there exists
	\begin{equation*}
	0 < p_{\min} \leq p(x)~~\textrm{for all $x \in \Xset$}
	\end{equation*}
	then the following statement holds: there exist constants $c$ and $C$ which depend only on $p_{\min},p_{\max}$ and $d$ such that if
	\begin{equation}
	\label{eqn:approximation_error}
	\norm{f}_{\Leb^2}^2 \geq C \cdot b^2 \cdot L^2 \cdot n^{-8/(8 + d)}
	\end{equation}
	then
	\begin{equation*}
	\norm{f}_n^2 \geq  2 \sqrt{6} b\sqrt{\frac{\kappa}{n^2}} + \frac{U_2(f;G_{n,r})}{4 n [\lambda_{\kappa}(G_{n,r})]^2}
	\end{equation*}
	with probability at least $1 - 5/b - 1/(nr^4) - C(n + r^{-d}) \exp\{-c nr^d\}$.
\end{lemma}
We note that for $r = n^{-2/(8 + d)}$, each of
\begin{equation*}
\frac{1}{nr^4}\to 0,~ C(n + r^{-d}) \exp\{-c nr^d\} \to 0~~\textrm{as}~ n \to \infty
\end{equation*}
and these contributions to the overall testing error are therefore asymptotically negligible. Theorem~\ref{thm:holder_testing_rate_2} then follows from combining Lemmas~\ref{lem:bias_correction_variance} and~\ref{lem:approximation_error} with Lemma~\ref{lem:fixed_graph_testing},

The following three Lemmas will be used in the proof of Lemma~\ref{lem:approximation_error}. Lemma~\ref{lem:empirical_norm_sobolev} -- copied from the graph testing notes -- implies a sufficient lower bound on  $\norm{f}_n^2$ when $s = 2$. It is stronger than we need, in the sense that it applies to all functions bounded in the second-order Sobolev norm rather than merely the second-order Holder norm; in our case we could obtain tighter bounds on the probability, but since we eventually hope to apply our argument to Sobolev functions anyway, we won't bother with this.

\begin{lemma}
	\label{lem:empirical_norm_sobolev}
	Let $\Xset$ be a Lipschitz domain over which the density is upper and lower bounded 
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty ~~\textrm{for all $x \in \Xset$,}
	\end{equation*}
	and let $f \in H^s(\Xset)$.Then for any $b \geq 1$, there exists $c_1$ such that if 
	\begin{equation}
	\norm{f}_{\Leb^2(\Xset)} \geq 
	\begin{cases*}
	c_1 \cdot b \cdot \norm{f}_{H^s(\Xset)} \cdot \max\{n^{-1/2},n^{-s/d}\},~~\textrm{if}~2s \neq d \\
	c_1 \cdot b \cdot \norm{f}_{H^s(\Xset)} \cdot n^{-a/2},~~\textrm{if}~ 2s = d ~\textrm{for any}~ 0 < a < 1
	\end{cases*}
	\end{equation}
	then,
	\begin{equation}
	\label{eqn:l2_to_empirical_norm_sobolev}
	\mathbb{P}\left[\norm{f}_n^2 \geq \frac{1}{b} \Ebb\bigl[\norm{f}_n^2\bigr]\right] \geq 1 - \frac{5}{b}
	\end{equation}
	where $c_1$ and $c_2$ are constants which may depend only on $s$, $\Xset$, $d$, $p_{\min}$ and $p_{\max}$.
\end{lemma}

Lemma~\ref{lem:neighborhood_eigenvalue} establishes that the eigenvalues $\lambda_{\kappa}(G_{n,r})$ scale at sufficiently fast rate as $\kappa$ increases.
\begin{lemma}
	\label{lem:neighborhood_eigenvalue}
	Suppose $p$ is bounded away from $0$ and $\infty$ everywhere on its support, i.e.
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty,~~ \textrm{for all $x \in [0,1]^d$.}
	\end{equation*}
	Then there exists constants $c$ and $C$ which depend only on $p_{\min}, p_{\max}$ and $d$, such that for all $1 \leq k \leq n$,
	\begin{equation}
	\label{eqn:neighborhood_eigenvalue}
	\lambda_k(G_{n,r}) \geq c \cdot \min\Bigl\{r^{d + 2}nk^{2/d}, nr^d\Bigr\} 
	\end{equation}
	with probability at least $1 - C (n + r^{-d}) \exp\bigl\{- c n r^d\bigr\}$. 
\end{lemma}

The first term on the right hand side of~\eqref{eqn:neighborhood_eigenvalue} is the scaling of the eigenvalues we desire. The second term is an artificial truncation on this spectrum. For the choices of $r$ and $\kappa$ in Lemma~\ref{lem:approximation_error}, these terms are exactly equal, and the truncation therefore does not come into play.

In Lemma~\ref{lem:graph_sobolev_seminorm}, we establish that $U$-statistic $U_2(f;G_{n,r})$ scales appropriately with the Holder norm of $f$.

\begin{lemma}
	\label{lem:graph_sobolev_seminorm}
	If $f \in C_0^2(\mathcal{X};L)$, $p \in C^1(\mathcal{X};L)$, and $n^{-1/d} \leq r \leq 1$, then there exists a constant $c$ such that
	\begin{equation*}
	\Bigl|U_2(f;G_{n,r})\Bigr| \leq c n^3 r^{2(d + 2)} L^2 p_{\max}^2
	\end{equation*}
	with probability at least $1 - 1/(nr^4)$.
\end{lemma}

\section{Proofs}

\subsection{Proof of Lemma~\ref{lem:fixed_graph_testing}}

In this proof, we will drop all notational dependence on the graph $G$ for ease of reading.

To prove Lemma~\ref{lem:fixed_graph_testing} we will first compute (bounds on) the expectation and variance of the test statistic $T_{\mathrm{spec}}$, and then use Chebyshev's inequality to show \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}.

\paragraph{Expectation of $T_{\mathrm{spec}}$:} Using the notation $\dotp{v}{w} = \sum_{i = 1}^{n} v_iw_i$, we have
\begin{align*}
\Ebb\bigl[T_{\spec}\bigr] & = \frac{1}{n}\sum_{k = 1}^{\kappa}\left( \dotp{\beta}{v_k}^2 + \Ebb\Bigl[ \dotp{\varepsilon}{v_k}^2 + 2 \dotp{\varepsilon}{v_k} \dotp{\beta}{v_k}\Bigr]\right) + \frac{1}{2 n\lambda_{\kappa}^2} \Ebb \Bigl[(\beta + \varepsilon)^T L (\beta + \varepsilon)\Bigr] \\
& = \frac{\kappa}{n} + \norm{\Pi_{\kappa,G}(\beta)}_n^2 + \frac{1}{2 n\lambda_{\kappa}^2}\biggl(\beta^T L \beta + \sum_{k = 1}^{n}\lambda_k\biggr)
\end{align*}

\paragraph{Variance of $T_{\mathrm{spec}}$:}
Note that $T_{\mathrm{spec}} = n^{-1} y^T V_{\kappa} V_{\kappa}^T y$ where $V_{\kappa}$ is the $n \times \kappa$ matrix with eigenvector $v_k$ as its $k$th column. Therefore, applying Cauchy-Schwarz,
\begin{align*}
\Var(T_{\spec}) & \leq \frac{2}{n^2} \biggl( \Var\Bigl(y^T V_{\kappa} V_{\kappa}^T y\Bigr) + \frac{1}{8 \lambda_{\kappa}^4}\Var\Bigl(y^T L y\Bigr) \biggr) \\
& = \frac{2}{n^2} \biggl( \Var\Bigl((\beta + \varepsilon)^T V_{\kappa} V_{\kappa}^T (\beta + \varepsilon)\Bigr) + \frac{1}{8 \lambda_{\kappa}^4}\Var\Bigl((\beta + \varepsilon)^T L (\beta + \varepsilon)\Bigr) \biggr) \\
& = \frac{2}{n^2}\biggl(4 n \norm{\Pi_{\kappa,G}(\beta)}_n^2 + 2 \kappa + \frac{1}{4\lambda_{\kappa}^4} \Bigl(2\beta^T L^2 \beta + \sum_{k = 1}^{n} \lambda_k^2 \Bigr)\biggr) \\
& \leq  \frac{2}{n^2}\biggl(4 n \norm{\Pi_{\kappa,G}(\beta)}_n^2 + 2 \kappa + \frac{1}{4\lambda_{\kappa}^4} \Bigl(2\beta^T L^2 \beta + n \lambda_n^2 \Bigr)\biggr) \\
& \leq \frac{2}{n^2}\biggl(4 n \norm{\Pi_{\kappa,G}(\beta)}_n^2 + 3 \kappa \biggr)
\end{align*}
where the last equality follows from standard properties of the Gaussian distribution, and the last inequality follows from~\eqref{eqn:graph_spectral_1}. We now move on to showing the desired inequalities \eqref{eqn:graph_spectral_type_I_error} and \eqref{eqn:graph_spectral_type_II_error}. Let $t(b) := b\sqrt{\frac{\kappa}{n^2}}$.

\paragraph{Proof of~\eqref{eqn:graph_spectral_type_I_error}:} By Chebyshev's inequality,
\begin{align*}
\Pbb_{\beta = 0}\biggl(T_{\spec} \geq \frac{\kappa}{n} + \frac{1}{2 n\lambda_{\kappa}^2}\sum_{k = 1}^{n} \lambda_k + 4t(b)\biggr)
& \leq \Pbb_{\beta = 0}\biggl(T_{\spec} - \Ebb\bigl[T_{\spec}\bigr] \geq \sqrt{6} t(b)\biggr) \\
& \leq \frac{\Var(T_{\spec})}{\bigl(\sqrt{6}t(b)\bigr)^2} = \frac{1}{b^2}.
\end{align*}

\paragraph{Proof of~\eqref{eqn:graph_spectral_type_II_error}:} For simplicity, we introduce the notation
\begin{equation*}
\Delta = \norm{\Pi_{\kappa,G}(\beta)}_n^2 + \frac{1}{2 n \lambda_{\kappa}^2} \beta^T L \beta
\end{equation*}
Assumption~\eqref{eqn:fixed_graph_testing_critical_radius_1} implies $\Delta \geq 2 \sqrt{6} \cdot t(b)$, and we have already shown that $\Ebb_{\beta}(T_{\spec}) = \Delta + \kappa/n + \bigl(\sum_{k = 1}^{n} \lambda_k\bigr)/(n\lambda_{\kappa}^2)$. Then another application of Chebyshev's inequality gives us
\begin{align*}
\Pbb_{\beta}\biggl(T_{\spec} \leq \frac{\kappa}{n} + \frac{1}{n\lambda_{\kappa}^2}\sum_{k = 1}^{n} \lambda_k^2 + \sqrt{6} \cdot t(b)\biggr) & = \Pbb_{\beta}\biggl(T_{\spec} - \Ebb_{\beta}(T_{\spec}) \leq \sqrt{6} \cdot t(b) - \Delta \biggr) \\
& \leq \Pbb_{\beta}\biggl(\abs{T_{\spec} - \Ebb_{\beta}(T_{\spec})} \leq \Delta/2 \biggr) \tag{since $\Delta \geq 2 \sqrt{6} \cdot t(b)$}	\\
& \leq \frac{4\Var_{\beta}(T_{\spec})}{\Delta^2} \\
& \leq \frac{24\kappa/n^2 + 32\Delta/n}{\Delta^2}.
\end{align*}
We handle each summand in the numerator separately. For the first term, since $\Delta \geq 2 \sqrt{6} \cdot t(b)$, we have
\begin{equation}
\label{eqn:spectral_type_II_error_pf1}
\frac{24\kappa}{n^2\Delta^2} \leq \frac{1}{b^2}.
\end{equation}

For the second term we have
\begin{equation}
\label{eqn:spectral_type_II_error_pf2}
\frac{32}{n\Delta} \leq \frac{16}{\sqrt{6} \cdot nt(b)} = \frac{16}{\sqrt{6} \cdot b\sqrt{\kappa}}, 
\end{equation}
and combining~\eqref{eqn:spectral_type_II_error_pf1} and~\eqref{eqn:spectral_type_II_error_pf2} yields~\eqref{eqn:graph_spectral_type_II_error}.

\paragraph{\eqref{eqn:fixed_graph_testing_critical_radius} implies~\eqref{eqn:fixed_graph_testing_critical_radius_1}.}

Since the eigenvalues $\lambda_k$ are ordered from smallest to largest, 
\begin{align*}
\norm{\Pi_{\kappa,G}(\beta)}_n^2 & = \norm{\beta}_n^2 - \frac{1}{n}\sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \\
& \geq \norm{\beta}_n^2 - \frac{1}{n \lambda_{\kappa}^2}\sum_{k = \kappa + 1}^{n} \dotp{\beta}{v_k}^2 \lambda_k^2 \\
& \geq \norm{\beta}_n^2 - \frac{\beta^T L^2 \beta}{n\lambda_{\kappa}^2},
\end{align*}
Therefore since we may rewrite the second-order graph Sobolev seminorm as,
\begin{align*}
\frac{\beta^T L^2 \beta}{n \lambda_{\kappa}^2} & = \frac{1}{n \lambda_{\kappa}^2} \sum_{i = 1}^{n} (L\beta)_i^2 \\
& = \frac{1}{4 n \lambda_{\kappa}^2} \sum_{i \in [n]^3} \Bigl(\beta_{i_1} - \beta_{i_2}\Bigr) \Bigl(\beta_{i_1} - \beta_{i_3}\Bigr) A_{i_1,i_2} A_{i_1,i_3} \\
& = \frac{1}{2 n \lambda_{\kappa}^2} \biggl(\beta^T L \beta + \frac{1}{2} \sum_{i \in (n)^3} \Bigl(\beta_{i_1} - \beta_{i_2}\Bigr) \Bigl(\beta_{i_1} - \beta_{i_3}\Bigr) A_{i_1,i_2} A_{i_1,i_3} \biggr) \\
& = \frac{1}{2 n \lambda_{\kappa}^2} \biggl(\beta^T L \beta + \frac{1}{2}U_2(\beta;G) \biggr)
\end{align*}
we have
\begin{equation*}
\norm{\Pi_{\kappa,G}(\beta)}_n^2 + \frac{\beta^T L \beta}{2 n \lambda_{\kappa}^2} \geq \norm{\beta}_n^2 - \frac{1}{4 n \lambda_{\kappa}^2}U_2(\beta;G)
\end{equation*}
whence we see that \eqref{eqn:fixed_graph_testing_critical_radius} implies~\eqref{eqn:fixed_graph_testing_critical_radius_1}.

\subsection{Proof of Lemma~\ref{lem:bias_correction_variance}}

By Lemma~\ref{lem:neighborhood_eigenvalue} there exist constants $c$ and $C$ such that 
\begin{equation}
\label{eqn:bias_correction_variance_pf2}
\lambda_{\kappa}(G_{n,r}) \geq c \cdot \min\Bigl\{n r^{d + 2} \kappa^{2/d}, n r^d\Bigr\}
\end{equation}
with probability at least $1 - C(n + r^{-d}) \exp\{-c nr^d\}$. Note that when $r = n^{-2/(8 + d)}$, and $\kappa = n^{2d/(8 + d)}$ as specified in the statement of Lemma~\ref{lem:bias_correction_variance}, the two terms inside the minimum are equal. We will now supply upper bounds on both $f^T L_{n,r}^2 f$ and $n \bigl(\lambda_n(G_{n,r})\bigr)^2$, which combined with~\eqref{eqn:bias_correction_variance_pf2} are sufficient to prove the claim of Lemma~\ref{lem:bias_correction_variance}.

\paragraph{Upper bound on $f^T L_{n,r}^2 f$.}
We begin by rewriting the quadratic form $f^T L_{n,r}^2 f$ as
\begin{equation*}
f^T L_{n,r}^2 f = 2 f^T L_{n,r} f + U_2(f;G_{n,r})
\end{equation*}
To upper bound the first term on the right hand side, we compute its expectation
\begin{align*}
\Ebb \Bigl[f^T L_{n,r} f\Bigr] & = n(n - 1) \Ebb\Bigl[\bigl(f(x_i) - f(x_j)\bigr)^2 K_r(x_i,x_j)\Bigr] \\
& \leq n(n - 1) L^2 r^{2} \Ebb\Bigl[K_r(x_i,x_j)\Bigr] \\
& \leq n(n - 1) L^2 r^{2 + d} p_{\max}.
\end{align*}
Since $f^T L_{n,r} f$ is always non-negative, by Markov's inequality
\begin{equation*}
f^T L_{n,r} f \leq b n(n - 1) L^2 r^{2 + d} p_{\max} 
\end{equation*}
with probability at least $1 - 1/b$.

In combination with the upper bound on $\abs{U_2(f;G_{n,r})}$ given by  Lemma~\ref{lem:graph_sobolev_seminorm}, choosing $c$ as in that Lemma we have
\begin{equation}
\label{eqn:bias_correction_variance_pf1}
f^T L_{n,r}^2 f \leq 2 b n^2 L^2 r^{2 + d} p_{\max} + c n^3 r^{2(d + 2)} L^2 p_{\max}^2
\end{equation}
with probability at least $1 - 1/b - 1/(nr^4)$.

Assuming the high probability bounds~\eqref{eqn:bias_correction_variance_pf2} and~\eqref{eqn:bias_correction_variance_pf1} hold, the routine calculations
\begin{equation*}
\frac{\Bigl(\lambda_{\kappa}(G_{n,r})\Bigr)^4 \kappa}{n^3 r^{2(d + 2)}} \geq n^3 r^{2(d + 2)} \geq n^{3} n^{-4(d + 2)/(d + 8)} = n^{(16 - d)/(8 + d)} \geq n^{1/2}
\end{equation*}
and
\begin{equation*}
\frac{\Bigl(\lambda_{\kappa}(G_{n,r})\Bigr)^4\kappa}{n^2 r^{(d + 2)}} \geq n^4 r^{3(d + 2)} \geq n^{4} n^{-6(d + 2)/(d + 8)} = n^{(20 - 2d)/(8 + d)} \geq n^{1/4} 
\end{equation*}
imply $f^T L_{n,r}^2 f \leq \kappa\Bigl(\lambda_{\kappa}(G_{n,r})\Bigr)^4$ for all $n$ sufficiently large.

\paragraph{Upper bound on $n \lambda_n^2$}
It is well known that for any graph $G$, it's largest eigenvalue $\lambda_n(G) \leq 2 \deg_{\max}(G)$. For the particular case of $G = G_{n,r}$, by Lemma~\ref{lem:max_degree}
\begin{equation*}
\deg_{\max}(G_{n,r}) \leq 2 p_{\max} \nu_d n r^d
\end{equation*}
with probability at least $1 -  n \exp\bigl\{- p_{\min} \nu_d n r^d \bigr\}$. Along with~\eqref{eqn:bias_correction_variance_pf2} this implies that for an appropriate choice of constant $c$
\begin{equation*}
\frac{\kappa \Bigl(\lambda_{\kappa}(G_{n,r})\Bigr)^4}{n \Bigl(\lambda_{n}(G_{n,r})\Bigr)^2} \geq c n r^{2d} \kappa = c n^{(8 - d)/(8 + d)}
\end{equation*}
whence the claim~\eqref{eqn:bias_correction_variance} follows since $d < 8$.

\subsection{Proof of Lemma~\ref{lem:approximation_error}}
Let us take the high probability conclusions of Lemmas~\ref{lem:empirical_norm_sobolev}, \ref{lem:neighborhood_eigenvalue}, and \ref{lem:graph_sobolev_seminorm} as given, namely that for $\kappa = n^{2d/(8 + d)}$ and $r = n^{-2/(8 + d)}$,
\begin{enumerate}
	\item $\norm{f}_n^2 \geq \frac{1}{b} \norm{f}_{\Leb^2}^2$,
	\item for some constant $c_1$, $\bigl[\lambda_{\kappa}(G_{n,r})\bigr]^2 \geq c_1 n r^{d + 2} \kappa^{2/d}$, and
	\item for some constant $c_2$, $\abs{U_2(f;G_{n,r})} \leq c_2 n^3 r^{2(d + 2)} \norm{f}_{C^2(\Xset)}^2$.
\end{enumerate}
Then, for a sufficiently large choice of constant $C$ in~\eqref{eqn:approximation_error},
\begin{align*}
2 \sqrt{6} b\sqrt{\frac{\kappa}{n^2}} + \frac{U_2(f;G_{n,r})}{4 n [\lambda_{\kappa}(G_{n,r})]^2} & \leq \Bigl(2 \sqrt{6} b + \frac{c_2}{c_1}\Bigr) n^{-8/(8 + d)} \\
& \leq \frac{1}{b}\norm{f}_{\Leb^2}^2 \\
& \leq \norm{f}_n^2.
\end{align*}
The probability with which this holds then comes from accumulating probabilities with which the conclusions of Lemmas~\ref{lem:empirical_norm_sobolev}, \ref{lem:neighborhood_eigenvalue}, and \ref{lem:graph_sobolev_seminorm} hold.

\subsection{Proof of Lemma~\ref{lem:neighborhood_eigenvalue}}
In the following proof, for graphs $G$ and $H$ defined on a common vertex set $V$, we write $G \preceq H$ if $f^T L_G f \leq f^T L_H f$ for every $f \in L^2(V)$.

We will prove Lemma~\ref{lem:neighborhood_eigenvalue} by comparing the graph $G_{n,r}$ to a graph $\wt{G}_{n,r}$ which we now define. Let $\wt{r} = r/(1 + \sqrt{d})$, and $M = 1/\wt{r}$; without loss of generality we will assume $M$ is an integer. The grid points
\begin{equation*}
\wb{Z} := \Bigl\{ \frac{1}{2M}(2m_1 - 1,\ldots,2m_d - 1): m \in [M]^d \Bigr\}
\end{equation*}
induces a tesselation $\mathcal{Q}\bigl(\wb{Z}\bigr)$ of $[0,1]^d$, defined as 
\begin{equation*}
Q(x) = \Bigl[x_1 - \frac{1}{2M}, x_1 + \frac{1}{2M}\Bigr] \times  \ldots \times \Bigl[x_d - \frac{1}{2M}, x_d + \frac{1}{2M}\Bigr],~~ \mathcal{Q}\bigl(\wb{Z}\bigr) := \set{Q(\wb{z}): \wb{z} \in \wb{Z}}
\end{equation*}
In the graph $\wt{G}_{n,r} = (X,\wt{E})$, there is an edge between $x_i$ and $x_J$ if they are in the same or adjoining cubes in $\mathcal{Q}(\wb{Z})$; formally speaking $(x_i,x_j) \in \wt{E}$ if either (a) there exists $\wb{z} \in \wb{Z}$ such that $x_i \in \mathcal{Q}(\wb{Z})$ and $x_j \in \mathcal{Q}(\wb{z})$ or (b) there exist $\wb{z}, \wb{z}' \in \mathcal{Q}(\wb{z}')$ such that $x_i \in \mathcal{Q}(\wb{z})$, $x_j \in \mathcal{Q}(\wb{z}')$, and $\norm{\wb{z} - \wb{z}} = 1/M$.

Let 
\begin{equation*}
Q_{\min} = \min_{\wb{z} \in \wb{Z}}~ \Bigl|Q(\wb{z}) \cap X\Bigr|,~~ Q_{\max} = \max_{\wb{z} \in \wb{Z}}~ \Bigl|Q(\wb{z}) \cap X\Bigr|
\end{equation*}

We will prove the following three claims. The first two are deterministic statements regarding graph comparisons, while the third involves the concentration of the functionals $Q_{\min}$ and $Q_{\max}$.  
\begin{enumerate}
	\item The graphs $G_{n,r}$ and $\wt{G}_{n,r}$ satisfy the partial ordering
	\begin{equation*}
	\wt{G}_{n,r}\preceq G_{n,r} 
	\end{equation*}
	\item For each $k = 1,\ldots,n$, the eigenvalues $\lambda_k(\wt{G}_{n,r})$ satisfy
	\begin{equation*}
	\lambda_k(\wt{G}_{n,r}) \geq \frac{Q_{\min}^2}{2 Q_{\max}}\min\Bigl\{\frac{k^{2/d}}{M^2}, d\Bigr\}
	\end{equation*}
	\item When $n$ is sufficiently large, there exist constants $c_1$ and $c_2$ such that
	\begin{equation*}
	c_1 p_{\min} n r^d \leq Q_{\min} \leq Q_{\max} \leq c_2 p_{\max} n r^d
	\end{equation*}
	with probability at least $1 - c_2 r^{-d} \exp \{-c_1 p_{\min} n r^d\}$. 
\end{enumerate}
The conclusion of Lemma~\ref{lem:neighborhood_eigenvalue} follows straightforwardly once these statements are proved.

\paragraph{Claim 1: Partial Ordering.}
By definition if $x_i \sim x_j$ in $\wt{G}_{n,r}$, then there exists $\wb{z}_{\ell}$ and $\wb{z}_{m}$ in $\wb{Z}$ such that
\begin{equation*}
\norm{x_i - \wb{z}_{\ell}}_2 \leq \frac{\sqrt{d}}{2M}, \norm{x_j - \wb{z}_{m}}_2 \leq \frac{\sqrt{d}}{2M}, \norm{\wb{z}_{\ell} - \wb{z}_m}_2 \leq \frac{1}{M}
\end{equation*}
Therefore by the triangle inequality,
\begin{equation*}
\norm{x_i - x_j}_2 \leq \frac{1 + \sqrt{d}}{M} = r
\end{equation*}
and so $x_i \sim x_j$ in $G_{n,r}$.

\paragraph{Claim 2: Eigenvalues of $\wt{G}_{n,r}$.}
Let $\wb{G}_d^{M}$ be the lattice graph formed over the grid points $\wb{Z}$, let $V_m = X \cap Q(\wb{z}_m)$ for each $m \in [M]^d$, and let $\wt{H} = (V_{\wt{H}}, E_{\wt{H}})$ where
\begin{equation*}
V_{\wt{H}} = \bigcup_{m \in [M]^d} \bigcup_{x_i \in V_m} (\wb{z}_m,x_i),~~ \textrm{and}~ \Bigl((\wb{z}_m,x_i),(\wb{z}_o,x_{j})\Bigr) \in E_{\wt{H}}~\textrm{if either $m = o$ or $z_m \sim z_o$ in $\wb{G}_d^{M}$} 
\end{equation*}
Lemma~\ref{lem:alden_product_graph} implies the following lower bound on the eigenvalues of $H$
\begin{align*}
\lambda_k(\wt{H}) & \geq \frac{1}{2} \frac{Q_{\min}^2}{Q_{\max}}\min\Bigl\{\lambda_k\bigl(\wb{G}_d^{M}\bigr),\deg_{\min}\bigl(\wb{G}_d^{M}\bigr)\Bigr\} \\
& \geq \frac{1}{2} \frac{Q_{\min}^2}{Q_{\max}}\min\Bigl\{\frac{k^{2/d}}{M},d\Bigr\}
\end{align*}
where the latter inequality follows from known facts about the eigenvalues and degrees of the $d$-dimensional lattice. Since $\wt{G}_{n,r}$ is isomorphic to $\wt{H}$, the same lower bound follows for the eigenvalues $\lambda_k(\wt{G}_{n,r})$.

\paragraph{Claim 3: Bounding $Q_{\min}$ and $Q_{\max}$.}
This follows from Hoeffding's inequality, as laid out in Lemma~\ref{lem:hoeffding_box}.

\subsection{Proof of Lemma~\ref{lem:graph_sobolev_seminorm}}
We will bound the expectation and variance of $U_2(f;G_{n,r})$; then prove the bound on the latter is sufficient to show concentration around the former at a sufficient rate.

For convenience, we introduce the notation 
\begin{equation*}
D_if(x) := (f(x_i) - f(x)) K_r(x_i,x),~~ K_r(x,z) := \1\Bigl\{\norm{x - z}_2 \leq r \Bigr\}
\end{equation*}

\paragraph{Expectation of $U_2(f;G_{n,r})$.}
By the linearity of expectation, and the law of iterated expectation
\begin{align*}
\Ebb\Bigl[U_2(f;G_{n,r})\Bigr] & = n (n - 1) (n - 2) \cdot \Ebb\Bigl[D_1f(x_2) D_1f(x_3)\Bigr] \\
& \leq n^3 \cdot \Ebb\biggl[ \Ebb\Bigl[D_1f(x_2)\bigm|x_1\Bigr]^2 \biggr].
\end{align*}
It therefore suffices to show 
\begin{equation}
\label{eqn:graph_sobolev_seminorm_pf1}
\biggl(\Ebb\Bigl[D_1f(x_2)\bigm|x_1\Bigr]\biggr)^2 \leq 2 p_{\max}^2 L^2 r^{2(2 + d)}
\end{equation}
we will do using Taylor expansions of $f$ and $p$. Writing the expectation as an integral, we have
\begin{equation*}
-\Ebb\Bigl[D_1f(x_2)\bigm|x_1 = x\Bigr] = \int \Bigl(f(y) - f(x)\Bigr)K_r(y,x)~ p(y) \,dy
\end{equation*}
Then, since $f \in C^2(\Xset;L)$, we have
\begin{equation*}
\abs{f(y) - f(x) - \bigl(\nabla f(x)\bigr)^T (y - x)} \leq L \norm{y - x}_2
\end{equation*}
for almost every $x \in \Xset$, and as a result
\begin{equation*}
\abs{\int \Bigl(f(y) - f(x)\Bigr)K_r(y,x)~ p(y) \,dy - \int \Bigl(\bigl(\nabla f(x)\bigr)^T (y - x)\Bigr)K_r(y,x)~ p(y) \,dy} \leq L p_{\max} r^{d + 2}
\end{equation*}
If $\mathrm{dist}(x,\partial \Xset) \leq r$, then since $f$ has first derivative zero at the boundary of $\Xset$, $\norm{\nabla f(x)}_2 \leq r$, which in turn gives
\begin{equation*}
\abs{\int \Bigl(\bigl(\nabla f(x)\bigr)^T (y - x)\Bigr)K_r(y,x)~ p(y) \,dy} \leq L p_{\max} r^{d + 2}
\end{equation*}
proving~\eqref{eqn:graph_sobolev_seminorm_pf1} when $\mathrm{dist}(x_1,\partial X) \leq r$.

Otherwise $\mathrm{dist}(x_1,\partial X) > r$. Since $p \in C^1(\Xset;p_{\max})$, we have
\begin{equation*}
\abs{p(y) - p(x)}  \leq p_{\max} r
\end{equation*}
and therefore
\begin{equation*}
\abs{\int \Bigl(\bigl(\nabla f(x)\bigr)^T (y - x)\Bigr)K_r(y,x)~ p(y) \,dy - p(x) \int \Bigl(\bigl(\nabla f(x)\bigr)^T (y - x)\Bigr)K_r(y,x) \,dy} \leq L p_{\max} r^{d + 2}.
\end{equation*}
However, since $K_r(y,x)$ is a radial kernel,
\begin{equation*}
\int \bigl(\nabla f(x)\bigr)^T (y - x) K_r(y,x)\,dy = 0
\end{equation*} 
and~\eqref{eqn:graph_sobolev_seminorm_pf1} follows by the triangle inequality.

\paragraph{Variance of $U_2(f;G_{n,r})$.}
We have
\begin{equation*}
\Var\Bigl(U_2(f;G_{n,r})\Bigr) = \sum_{i \in (n)^3} \sum_{j \in (n)^3} \Cov\Bigl(D_{i_1}f(x_{i_2}) \cdot D_{i_1}f(x_{i_3}),D_{i_4}f(x_{i_5})\cdot D_{i_4}f(x_{i_5})\Bigr)
\end{equation*}
Let $I = i \cup j$ be the joint index set over $i$ and $j$. Our bound on the covariance term in the preceding display will be a function of the cardinality $\abs{I}$. For example, when $\abs{I} = 6$, by the independence properties of $X$ the covariance is equal to 0. Otherwise if $\abs{I} = 3,4$ or $5$, we have
\begin{align*}
\biggl|\Cov\Bigl(& D_{i_1}f(x_{i_2}) \cdot D_{i_1}f(x_{i_3}),D_{i_4}f(x_{i_5})\cdot D_{i_4}f(x_{i_5})\Bigr)\biggr| \\
& \leq \biggl|\Ebb\Bigl[D_{i_1}f(x_{i_2}) \cdot D_{i_1}f(x_{i_3}) \cdot D_{i_4}f(x_{i_5})\cdot D_{i_4}f(x_{i_5})\Big]\biggr| + \biggl(\Ebb\Bigl[D_{i_1}f(x_{i_2}) \cdot D_{i_1}f(x_{i_3})\Bigr] \biggr)^2 \\
& \leq \biggl|\Ebb\Bigl[D_{i_1}f(x_{i_2}) \cdot D_{i_1}f(x_{i_3}) \cdot D_{i_4}f(x_{i_5})\cdot D_{i_4}f(x_{i_5})\Big]\biggr| + p_{\max}^4 L^4 r^{4(d + 2)} \\
& \leq L^4 r^4 ~ \Pbb \biggl(G[X_I]~\textrm{is connected}\biggr) + 4 p_{\max}^4 L^4 r^{4(d + 2)} \\
& \leq L^4 r^4 p_{\max}^{(\abs{I} - 1)} (\nu_d r)^{d(\abs{I} - 1)} (\abs{I} - 1)! + 4 p_{\max}^4 L^4 r^{4(d + 2)} \\
& \leq L^4 p_{\max}^4 \Bigl(r^{d(\abs{I} - 1)} + r^{4(d + 2)}\Bigr)
\end{align*}
where $c_1 := 24 \nu_d^{4d} + 4$.

Our bound depends on the indices $I$ only through $\abs{I}$, and we reiterate that the covariance is non-zero only when $\abs{I} = 3,4$ or $5$. Using our previous expression on the variance of the U-statistic, we see
\begin{align*}
\Var\Bigl(U_2(f;G_{n,r})\Bigr) & \leq c_1 L^4 p_{\max}^4 \sum_{\abs{I} = 3}^{5} n^{\abs{I}} \Bigl( r^{4 + d(\abs{I} - 1)} + r^{4(d + 2)} \Bigr) \\
& \leq 3 c_1 L^4 p_{\max}^4 n^{5} r^{4 + 4d}
\end{align*}
where the latter inequality follows since $n^{-1/d} \leq r \leq 1$.

\paragraph{Concentration.}
Now we simply apply Chebyshev's inequality to deduce
\begin{equation*}
\Pbb \Bigl(\abs{U_2(f;G_{n,r}) - \Ebb(U_2(f;G_{n,r}))} \geq \sqrt{3c_1} L^2 p_{\max}^2 n^3 r^{2(2+d)}\Bigr) \leq \frac{1}{nr^4}.
\end{equation*}
Therefore by the triangle inequality,
\begin{equation*}
\abs{U_2(f;G_{n,r})} \leq (\sqrt{3c_1} + 2) p_{\max}^2 L^2 n^3 r^{2(2+d)}
\end{equation*}
with probability at least $1 - (nr^4)^{-1}$. This completes the proof of Lemma~\ref{lem:graph_sobolev_seminorm}, upon proper choice of the constant $c = \sqrt{3c_1} + 2$.

\section{Additional Theory}

\subsection{Concentration}

The maximum eigenvalue of a neighborhood graph is no greater than two times its degree, which can be upper bounded with high probability using, for example, Hoeffding's Inequality.
\begin{lemma}
	\label{lem:max_degree}
	Suppose the density $p$ is bounded away from zero and infinity over its support
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty,~~\textrm{for all $x \in [0,1]^d$.}
	\end{equation*}
	Then the maximum degree of $G_{n,r}$ can be upper bounded
	\begin{equation*}
	\deg_{\max}(G_{n,r}) \leq 2 p_{\max} \nu_d n r^d
	\end{equation*} 
	with probability at least $1 - n \exp\Bigl\{- p_{\min} \nu_d n r^d \Bigr\}$. 
\end{lemma}

Similarly, Hoeffding's inequality can be used to upper (and lower) bound the maximum (and minimum) number of points in any cube $Q(\wb{z})$. 
\begin{lemma}
	\label{lem:hoeffding_box}
	Suppose the density $p$ is bounded away from zero and infinity over its support
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty,~~\textrm{for all $x \in [0,1]^d$.}
	\end{equation*}
	Then the maximum and minimum number of samples in any grid cell can be bounded,
	\begin{equation*}
	Q_{\min} \geq \frac{p_{\min}}{2 (1 + \sqrt{d})^d} n r^d
	\end{equation*}
	and
	\begin{equation*}
	Q_{\max} \leq \frac{2 p_{\max}}{(1 + \sqrt{d})^d} n r^d
	\end{equation*} 
	with probability at least $1 - \frac{2(1 + \sqrt{d})^d}{r^d} \exp\Bigl\{-\bigl(p_{\min} (1 + \sqrt{d})^d/8 \bigr)nr^d \Bigr\}$. 
\end{lemma}

\subsection{Graph Comparison}

For an integer $\mc{M} \geq 1$, we introduce the \textcolor{red}{(Alden product)} graph. Let $G = (Z,E)$ with $\abs{Z} = \{z_1,\ldots,z_{\mc{M}}\}$, and let $V_1,\ldots,V_{\mc{M}}$ be sets of size $N_1,\ldots,N_{\mc{M}}$. The \textcolor{red}{(Alden product)} graph $H$ is defined as
\begin{equation*}
H = \Bigl(V_H, E_H\Bigr)
\end{equation*}
where $V_H = \bigcup_{m = 1}^{\mc{M}} \bigcup_{j = 1}^{N_m} (z_m,x_{m,j})$ and  $(z_{\ell},x_{\ell,j}) \sim (z_m,x_{m,j})$ in $H$ if (a) $z_{\ell} = z_m$ or (b) $z_{\ell} \sim z_m$ in $G$. Let
\begin{equation*}
N_{\min} := \min_{m \in [\mc{M}]} N_m,~~ N_{\max} := \max_{m \in [\mc{M}]} N_m
\end{equation*} 
and
\begin{equation*}
\deg_{\min}(G) := \min_{m \in [\mc{M}]} \deg(z_m;G),~~ \deg_{\max}(G) := \max_{m \in [\mc{M}]} \deg(z_m;G)
\end{equation*}
\begin{lemma}
	\label{lem:alden_product_graph}
	Assume that $G$ is a non-empty graph, and that for each $m \in [\mc{M}]$ the set $V_m$ is non-empty set. 
	\begin{enumerate}
		\item If $k \in [\mc{M}]$, 
		\begin{equation*}
		\frac{1}{2}\frac{N_{\min}^2}{N_{\max}} \lambda_k(G) \leq \lambda_k(H) \leq \frac{N_{\max}^2}{N_{\min}} \lambda_k(G)
		\end{equation*}
		\item Otherwise if $k = \mc{M} + 1,\ldots,\sum_{m = 1}^{\mc{M}}N_m$,
		\begin{equation*}
		\frac{N_{\min}^2}{N_{\max}} \deg_{\min}(G) \leq \lambda_k(H) \leq 2 \frac{N_{\max}^2}{N_{\min}} \deg_{\max}(G)
		\end{equation*}
	\end{enumerate}
\end{lemma}
\begin{proof}
	To prove Lemma~\ref{lem:alden_product_graph}, we will 
	\begin{enumerate}
		\item Exhibit an orthonormal basis of $L^2(V_H)$, then
		\item Compute upper and lower bounds on the Rayleigh quotient, then
		\item Apply the Courant-Fischer Theorem, and then
		\item Put together the pieces.
	\end{enumerate}

\paragraph{Step 1: Orthonormal basis of $L^2(V_H)$.}
Let $v_1,\ldots,v_{\mc{M}}$ be eigenvectors of $L_G$, with corresponding eigenvalues $\lambda_1(G) \leq \cdots \leq \lambda_{\mc{M}}(G)$, Then for $k = 1,\ldots,\mc{M}$, let
\begin{equation*}
v_k^{(H)}\bigl((z_m,x_{m,j})\bigr) = \frac{1}{\sqrt{N_m}} v_k(z_m) 
\end{equation*}
Now, for each $m = 1,\ldots,\mc{M}$, there is an $(N_m - 1)$-dimensional subspace of $L^2(V_m)$ orthogonal to the constant function. Let $u_{1,m}, \ldots, u_{N_m - 1,m}$ be an orthonormal sequence spanning this subspace, and further let
\begin{equation*}
u_{\ell,m}^{(H)}\bigl((z_{m'},x_{m',j})\bigr) = 
\begin{cases*}
u_{\ell,m}(x_{m,j}),&~~\textrm{if $m = m'$}\\
0,&~~\textrm{otherwise.}
\end{cases*}
\end{equation*}
for $m \in [\mc{M}]$ and $\ell \in [N_m - 1]$.

We now verify that $\{v_k^{(H)}: k \in [\mc{M}]\} \cup \{u_{\ell,m}^{(H)}: m \in [\mc{M}], \ell \in [N_m - 1] \}$ is an orthonormal basis of $L^2(V_H)$. First,
\begin{align*}
\dotp{v_k^{(H)}}{v_{\ell}^{(H)}}_{L^2(V_H)} & = \sum_{m = 1}^{\mc{M}} \sum_{j = 1}^{N_m} v_k^{(H)}\bigl((z_m,x_{m,j})\bigr) v_{\ell}^{(H)}\bigl((z_m,x_{m,j})\bigr) \\
& = \sum_{m = 1}^{\mc{M}} \sum_{j = 1}^{N_m} \frac{v_k(z_m) v_{\ell}(z_m)}{N_m} \\
& = \sum_{m = 1}^{\mc{M}} v_k(z_m) v_{\ell}(z_m) \\
& = \1\{k = \ell\}
\end{align*} 
Next,
\begin{align*}
\dotp{u_{\ell,m}^{(H)}}{u_{\ell',m'}^{(H)}}_{L^2(V_H)} & = \sum_{o = 1}^{\mc{M}} \sum_{j = 1}^{N_o} u_{\ell,m}(x_{o,j}) u_{\ell',m'}(x_{o,j}) \1\{m = o\} \1\{m' = o\} \\
& = \sum_{j = 1}^{N_m} u_{\ell,m}(x_{m,j}) u_{\ell',m}(x_{m,j}) \1\{m = m'\} \\
& = \1\{m = m'\} \1\{\ell = \ell'\}
\end{align*}
Finally,
\begin{align*}
\dotp{v_k^{(H)}}{u_{\ell,m}^{(H)}}_{L^2(V_H)} & = \sum_{o = 1}^{\mc{M}} \sum_{j = 1}^{N_o} \frac{v_k(z_o)}{\sqrt{N_o}} u_{\ell,m}(x_{o,j}) \1\{m = o\} \\
& = \frac{v_k(z_m)}{\sqrt{N_m}} \sum_{j = 1}^{N_o} u_{\ell,m}(x_{m,j}) \\
& = 0,
\end{align*}
where the last equality follows since $u_{\ell,m}$ is orthogonal to the constant function in $L^2(V_m)$. Since $\dim L^2(V_H) = \sum_{m = 1}^{\mc{M}} N_m$ and we have exhibited a total of $\sum_{m = 1}^{\mc{M}} N_m$ orthonormal functions in $L^2(V_H)$, they form a basis.

\paragraph{Step 2: Bounding the Rayleigh quotient.}

Let $N: L^2(V_H) \to L^2(V_H)$ be the following diagonal operator,
\begin{equation*}
\Bigl(N v\Bigr) \Bigl((z_m,x_{m,j})\Bigr) = N_m v\bigl((z_m,x_{m,j})\bigr)
\end{equation*}
and let $\wt{L}_H = N^{1/2} L_H N^{1/2}$. We will analyze the Rayleigh quotient of the operator $\wt{L}_H$ with respect to the basis $\bigl\{v_k^{(H)}, u_{\ell,m}^{(H)}\bigr\}$.

We begin by establishing the following upper and lower bounds,
\begin{equation}
\label{eqn:alden_product_graph_pf1}
N_{\min}^2 \lambda_k(G) \leq \dotp{\wt{L}_H v_k^{(H)}}{v_k^{(H)}}_{L^2(V_H)} \leq N_{\max}^2 \lambda_k(G). 
\end{equation}
Seeing as $N^{1/2}$ is self-adjoint, 
\begin{align*}
\Bigl \langle\wt{L}_H v_k^{(H)}, v_k^{(H)} \Bigr \rangle_{L^2(V_H)} & = \Bigl \langle L_H N^{1/2} v_k^{(H)},N^{1/2}v_k^{(H)} \Bigr \rangle_{L^2(V_H)} \\ 
& = \frac{1}{2}\sum_{m = 1}^{\mc{M}} \sum_{o = 1}^{\mc{M}} \sum_{i = 1}^{N_m} \sum_{j = 1}^{N_o} \Bigl(v_k(z_m) - v_k(z_o)\Bigr)^2 \1\{z_m \sim z_o~\textrm{in $G$}\} \\
& = \frac{1}{2}\sum_{m = 1}^{\mc{M}} \sum_{o = 1}^{\mc{M}} N_m N_o \Bigl(v_k(z_m) - v_k(z_o)\Bigr)^2 \1\{z_m \sim z_o~\textrm{in $G$}\} \\
& \leq N_{\max}^2 \lambda_k(G),
\end{align*}
and by reversing the last inequality we have
\begin{equation*}
\dotp{\wt{L}_H v_k^{(H)}}{v_k^{(H)}}_{L^2(V_H)} \geq N_{\min}^2 \lambda_k(G).
\end{equation*}

Next we show that for all $m \in [\mc{M}], \ell \in [N_m - 1]$,
\begin{equation}
\label{eqn:alden_product_graph_pf2}
N_{\min}^2\Bigl(\deg_{\min}(G) + 1\Bigr) \leq \dotp{\wt{L}_H u_{\ell,m}^{(H)}}{u_{\ell,m}^{(H)}}_{L^2(V_H)} \leq N_{\max}^2 \Bigl(\deg_{\max}(G) + 1\Bigr) 
\end{equation}
Expanding the inner product as double sum, we have
\begin{align}
\Bigl \langle & \wt{L}_H u_{\ell,m}^{(H)}, u_{\ell,m}^{(H)} \Bigr \rangle_{L^2(V_H)} \\
& = \frac{1}{2} \sum_{m' = 1}^{\mc{M}} \sum_{o = 1}^{\mc{M}} \sum_{i = 1}^{N_{m'}} \sum_{j = 1}^{N_o} \Bigl( \sqrt{N_{m'}} u_{\ell,m}(x_{m,i}) \1\{m' = m\} - \sqrt{N_o} u_{\ell,m}(x_{m,i}) \1\{o = m\}\Bigr)^2 \1\{m' = o~\textrm{or}~z_{m'} \sim z_o~\textrm{in $G$}\} \nonumber \\
& =  \frac{N_m}{2}\biggl(\underbrace{\sum_{i,j = 1}^{N_m} \Bigl( u_{\ell,m}(x_{m,i}) -  u_{\ell,m}(x_{m,j})\Bigr)^2}_{:=S_1} + 2 \underbrace{\sum_{o = 1}^{\mc{M}} \sum_{i = 1}^{N_{m}} \sum_{j = 1}^{N_o}  \Bigl( u_{\ell,m}(x_{m,i})\Bigr)^2 \1\{z_m \sim z_o~\textrm{in $G$}\}}_{:=S_2} \biggr) \label{eqn:alden_product_graph_pf3}
\end{align}
$S_1$ is the sum over all term where $m' = m = o$, and $S_2$ is the sum over all terms where either $m' = m \neq o$ or $m' \neq m = o$. Then, since by definition $u_{\ell,m}$ is orthonormal to the constant function and has unit norm in $L^2(V_m)$, we have
\begin{equation*}
2N_{\min} \leq S_1 = \sum_{i,j = 1}^{N_m} \Bigl( \bigl(u_{\ell,m}(x_{m,i})\bigr)^2 + \bigl(u_{\ell,m}(x_{m,j})\bigr)^2 - u_{\ell,m}(x_{m,i}) u_{\ell,m}(x_{m,j})\Bigr) = 2N_m \leq 2N_{\max}
\end{equation*}
and
\begin{equation*}
N_{\min} \deg_{\min}(G) \leq S_2 = \biggl(\sum_{i = 1}^{N_{m}} \Bigl(u_{\ell,m}(x_{m,i})\Bigr)^2 \biggr) \biggl(\sum_{o = 1}^{\mc{M}} N_o\biggr) \leq N_{\max} \deg_{\max}(G)
\end{equation*}
These bounds along with~\eqref{eqn:alden_product_graph_pf3} imply~\eqref{eqn:alden_product_graph_pf2},

\paragraph{Step 3: Courant-Fischer Theorem.}

Now we translate the upper and lower bounds we established in Step 2 into upper and lower bounds on $\wt{\lambda}_k(H) = \lambda_k(\wt{L}_H)$. Naturally, we do this by means of the Courant-Fischer Theorem, which implies
\begin{align*}
\wt{\lambda}_k(H) & = \min_{V}~ \Bigl\{\max_{v \in V} ~\bigl \langle \wt{L}_Hv , v \bigr \rangle \Bigr\} \\
\wt{\lambda}_k(H) & = \max_{V}~ \Bigl\{\min_{v \in V^{\perp}}~ \bigl \langle \wt{L}_Hv , v \bigr \rangle \Bigr\}
\end{align*}
where the outer minimum (maximum) are taken over all subspaces $V$ of dimension $k$. We now construct subspaces $V_k$ for $k = 1,\ldots, \sum_{m = 1}^{\mc{M}} N_m$. If $k \in [\mc{M}]$, let $V_k = \mathrm{span}\{v_1^{(H)}, \ldots, v_k^{(H)}\}$; otherwise if $k > \mc{M}$, let $V_k = V_{\mc{M}} \cup U_{k - \mc{M}}$, where $U_{k - \mc{M}}$ is the span of $(k - \mc{M})$-functions $u_{\ell,m}$ chosen arbitrarily. The following facts are immediate consequences of Courant-Fischer and our derivations in Step 2:
\begin{enumerate}[(a)]
	\item For each $k \in [\mc{M}]$,
	\begin{equation*}
	\wt{\lambda}_k(H) \leq \max_{v \in V_k} ~\bigl \langle \wt{L}_Hv , v \bigr \rangle \leq \lambda_k(G) N_{\max}^2 
	\end{equation*}
	\item For each $k > \mc{M}$,
	\begin{equation*}
	\wt{\lambda}_k(H) \leq \max_{v \in V_k} ~\bigl \langle \wt{L}_Hv , v \bigr \rangle \leq \max \biggl\{ \lambda_{\mc{M}}(G) N_{\max}^2, N_{\max}^2 \Bigl(\deg_{\max}(G) + 1\Bigr)  \biggr\} \leq 2 \deg_{\max}(G) N_{\max}^2
	\end{equation*}
	\item For each $k \in [\mc{M}]$,
	\begin{equation*}
	\wt{\lambda}_k(H) \geq \min_{v \in V_k^{\perp}} ~\bigl \langle \wt{L}_Hv , v \bigr \rangle \geq \min\biggl\{\lambda_k(G) N_{\min}^2 , N_{\min}^2 \cdot \Bigl(\deg_{\min}(G) + 1\Bigr) \biggr\} \geq \frac{1}{2}\lambda_k(G) N_{\min}^2
	\end{equation*}
	\item For each $k > \mc{M}$,
	\begin{equation*}
	\wt{\lambda}_k(H) \geq \min_{v \in V_k^{\perp}} ~\bigl \langle \wt{L}_Hv , v \bigr \rangle \geq N_{\min}^2 \Bigl(\deg_{\min}(G) + 1\Bigr)
	\end{equation*}
\end{enumerate}

\paragraph{Step 4: Completing the Proof.}
Let $v$ be an eigenvector of $\wt{L}_H$, so that
\begin{equation*}
\wt{L}_H v = \lambda v.
\end{equation*}
Then, 
\begin{equation*}
w = \frac{N^{1/2}v}{\norm{N^{1/2}v}_{L^2(V_H)}}
\end{equation*}
is a solution to the generalized eigenvalue equation
\begin{equation*}
L_H w = \lambda N^{-1} w.
\end{equation*}
(To see this, simply left multiple both sides of the generalized eigenvalue equation by $N^{1/2}$. Also note that $N^{-1}$ is positive semi-definite since by assumption $N_m > 0$ for all $m$, and thus invertible.)  Therefore,
\begin{equation*}
\frac{\lambda}{N_{\max}} \leq \dotp{L_H w}{w}_{L^2(V_H)} \leq \frac{\lambda}{N_{\min}}
\end{equation*}
and using Courant-Fischer as in Step 3, we have that for each $k = 1,\ldots,\sum_{m = 1}^{\mc{M}} N_m$,
\begin{equation*}
\frac{\wt{\lambda}_k(H)}{N_{\max}} \leq \lambda_k(H) \leq \frac{\wt{\lambda}_k(H)}{N_{\min}}
\end{equation*}
Along with the bounds of Step 3, this establishes the claim of Lemma~\ref{lem:alden_product_graph}.

\end{proof}
\end{document}

### 2020-08-19 notes/weekly_notes/3.5.20.NotesForWeek.tex
Added weekly notes, which include some derivations of testing results for the fixed graph case.

@@
 For an arbitrary graph $G$ on $n$ nodes with graph Laplacian $L_G = D_G - A_G$, let $\Lambda(G)$ consist of the eigenvalues $0 = \lambda_1(G) \leq \lambda_2(G) \leq \cdots \leq \lambda_n(G)$ of $L_G$, with $v_k(G) = (v_{k,1}(G),\ldots,v_{k,n}(G)) \in \Reals^n$ denoting the eigenvector corresponding to the $i$th eigenvalues $\lambda_k(G)$.
-The graph Laplacian eigenvector projection test is a linear projection test; letting $\kappa$ be some integer between $1$ and $n$, we define $\Pi_{k,G}(\theta) := \frac{1}{\sqrt{n}}\sum_{k = 1}^{\kappa} \dotp{v_k}{\theta}$ to be a suitable rescaling of the projection of $\theta$ onto the span of $v_1(G),\ldots,v_k(G)$.
+The graph Laplacian eigenvector projection test is a linear projection test; letting $\kappa$ be some integer between $1$ and $n$, we define $\Pi_{k,G}(\theta) := \sum_{k = 1}^{\kappa} \dotp{v_k}{\theta}$ to be the projection of $\theta$ onto the span of $v_1(G),\ldots,v_k(G)$.
 
@@
 \begin{equation*}
-a_{k - 1} =
-\begin{cases*}
-(k - 1)^s,~~ \textrm{if $k - 1$ is even} \\
-(k - 2)^s,~~ \textrm{if $k - 1$ is odd}
-\end{cases*}
+a_{k} = (k - 1)^s
 \end{equation*}
-and for general $d$, $a_k = \sqrt{\sum_{j = 1}^{d} a_{k_j}^2}$. It is not hard to show that $\wt{H}^s([0,1]^d)$ consists of those functions in $H^s([0,1]^d)$ which satisfy some boundary conditions
+and for general $d$, $a_k = \sqrt{\sum_{j = 1}^{d} a_{k_j}^2}$.
+It is not hard to show that, when $s$ is an integer, $\wt{H}^s([0,1]^d)$ consists of those functions in $H^s([0,1]^d)$ which satisfy some Neumann-type boundary conditions.
+For example when $d = 1$,
 \begin{equation*}
-\wt{H}^s([0,1]^d;L) = \Bigl\{f \in H^{s}([0,1]^d;L'): \forall x_0 \in \partial [0,1]^d, \frac{\partial^{\ell} f}{\partial \mathbf{n}^{\ell}}(x_0) = 0,~~\textrm{for $\ell = 1,\ldots,s-1$} \Bigr \} \supseteq H_0^s([0,1]^d;L')
+\wt{H}^s([0,1];L) = \Bigl\{f \in H^{s}([0,1];L \pi^{s}): f^{(\ell)}(0) = f^{(\ell)}(1) = 0,~~\textrm{for all odd integers $0 < \ell < s$} \Bigr \} \supseteq H_0^s([0,1];L'),
 \end{equation*}
-for a constant $L'$ which depends only on $L$ and $s'$; we work through the details for the case $d = 1$ in Section~\ref{subsec:sobolev_class_equivalence}.
+a fact which we prove in Section~\ref{subsec:sobolev_class_equivalence}.
 
@@
 	\label{eqn:grid_sobolev_approximation_error_6}
-	\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,\wb{G}_d}(f)}_{n}^2 \leq c L^{1/2} \norm{f}_{\Leb^2} n^{-s/d} + \sum_{k \in \Nbb \setminus [K]^d} \theta_k^2
+	\norm{f}_{\Leb^2}^2 - \norm{\Pi_{\kappa,\wb{G}_d}(f)}_{n}^2 \leq c L^{1/2} \norm{f}_{\Leb^2} n^{-s/d} + \sum_{k \in \Nbb^d \setminus [K]^d} \theta_k^2
 	\end{equation}
@@
 
-\textcolor{red}{Work through details.}
+We will show that when $s \geq 1$,
+\begin{equation*}
+\wt{H}^s([0,1];L) = \Bigl\{f \in H^{s}([0,1];L'): f^{(\ell)}(0) = f^{(\ell)}(1) = 0,~~\textrm{for each $\ell < s$ odd.} \Bigr \}
+\end{equation*}
 
+First, we show the $\subseteq$ direction. Let $f = \sum_{k = 1}^{\infty} \theta_k \varphi_k$ satisfy $\sum_{k = 1}^{\infty} \theta_k^2 a_k^2 \leq L^2$. Put
+\begin{equation*}
+f_{\diamond} := \sqrt{2} \sum_{k = 1}^{\infty} \theta_k \phi_k
+\end{equation*}
+where the functions $\phi_k = \cos(\pi[k - 1]x)$ are orthonormal in $\Leb^2([-1,1])$.
+Since $\sum_{k = 1}^{\infty} (\sqrt{2}\theta_k)^2 a_k^2 \leq 2L^2$, we know $f_{\diamond} \in H_{\mathrm{per}}^s([-1,1];2L)$.
+Moreover since $f_{\diamond}$ is an even function, we have that
+\begin{equation*}
+\sum_{\ell = 0}^{s}\int_{0}^{1} \bigl(f_{\diamond}^{(\ell)}(x)\bigr)^2 \,dx = \frac{1}{2} \sum_{\ell = 0}^{s}\int_{-1}^{1} \bigl(f_{\diamond}^{(\ell)}(x)\bigr)^2 \,dx \leq \frac{1}{\pi^{2s}}L^2
+\end{equation*}
+where the latter inequality follows from \textcolor{red}{Tsybakov}.
+As $f(x) = f_{\diamond}(x)$ for all $x \in [0,1]$, we have established that $f \in H^s([0,1],L')$ for $L' = L/\pi^s$. As for the boundary cases $x = 0$ and $x = 1$, putting
+\begin{equation*}
+f_N := \sum_{k = 1}^{N} \theta_k \varphi_k
+\end{equation*}
+we observe the following facts: first that $f_N^{(\ell)}(0) = f_N^{(\ell)}(1) = 0$ for all $N \in \Nbb$ and $\ell$ odd, and second that $f_N^{(\ell)}$ converges uniformly for all $0 \leq \ell < s$.
+Therefore we have $f^{(\ell)}(x) = \lim_{N \to \infty} f_N^{(\ell)}(x)$ for all $x \in [0,1]$ and $0 \leq \ell < s$, and in particular $f^{(\ell)}(0) = f^{(\ell)}(1) = 0$ for all $0 < \ell < s$ odd.
+
+Now for the $\supseteq$ direction. Let $f \in H^s([0,1];L)$ satisfy $f^{(\ell)}(0)$ for all $\ell = 1,\ldots,s - 1$ odd. Put
+\begin{equation*}
+f_{\diamond} = f(\abs{x})~~\textrm{for $x \in [-1,1]$}
+\end{equation*}
+and note the following facts: first that $f_{\diamond}^{(\ell)}(0)$ is well-defined for each $\ell = 0,\ldots, s - 1$, and second that $f_{\diamond}^{(s - 1)}$ is absolutely continuous on $[-1,1]$.
+As a result, $f_{\diamond}$ is $s$-times weakly differentiable (and $s - 1$ times classically differentiable), and by construction satisfies
+\begin{equation*}
+\norm{f_{\diamond}}_{H^s([-1,1])} = \sqrt{2} \norm{f}_{H^s([0,1])} \leq 2L
+\end{equation*} 
+Since clearly $f_{\diamond}^{(\ell)}(-1) = f_{\diamond}^{(\ell)}(1)$ for each $\ell = 0,\ldots, s - 1$, we have that $f_{\diamond} \in H_{\mathrm{per}}^s([-1,1];\sqrt{2}L)$.
+Since $f_{\diamond}$ is even, it may be expressed as $f_{\diamond} = \sum_{k = 1}^{\infty} \theta_k \phi_k$ where $\sum_{k = 1}^{\infty} \theta_k^2 a_k^{2} \leq 2L^2/\pi^{2s}$.
+Since $f(x) = f_{\diamond}(x)$ for $x \in [0,1]$, this implies
+\begin{equation*}
+f(x) = \sum_{k = 1}^{\infty} \theta_k \phi_k(x) = \sum_{k = 1}^{\infty} \frac{1}{\sqrt{2}}\theta_k \varphi_k(x) 
+\end{equation*}
+and the claim is proved.
 \end{document}

### 2020-08-19 notes/weekly_notes/4.16.20.NotesForWeek.tex
Added weekly notes, which include some derivations of testing results for the fixed graph case.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\makeatletter
\newcommand{\vast}{\bBigg@{4}}
\newcommand{\Vast}{\bBigg@{5}}
\makeatother

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}
\newcommand{\Lap}{{\bf L}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\LE}{\mathrm{LE}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 4/10/20 - 4/16/20}
\author{Alden Green}
\date{\today}
\maketitle

Let $\mc{M}$ be a closed, connected, smooth manifold without boundary of dimension $m$ embedded in $\Rd$. We give to $\mc{M}$ the Riemannian structure induced by the ambient space $\Rd$, and denote the volume form by $\mu$.  Let $P$ be a distribution supported on $\mc{M}$, with density $p$ with respect to the volume form $\mu$. Suppose we observe $n$ samples $X_1, \ldots, X_n$ drawn independently from $P$. In addition, we observe responses
\begin{equation*}
Y_i = f_0(X_i) + \varepsilon_i
\end{equation*}
where $\varepsilon_i$ are i.i.d $N(0,1)$ noise, and $f_0:\mc{M} \to \Reals$ is an unknown function we wish to estimate.

We construct a neighborhood graph $G_{n,r}$ over the data $X_1,\ldots,X_n$ as follows. For a kernel $K: \Rd \to [0,\infty)$ and radius $r > 0$, we let
\begin{equation*}
K_r(X_i,X_j) = K\biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\biggr)
\end{equation*}
be the weight of the edge connecting $X_i$ and $X_j$. The graph $G_{n,r}$ has an associated Laplacian matrix $L_{n,r}$, which is defined by the action
\begin{equation*}
\Lap_{n,r}f(X_i) := \sum_{i = 1}^{n} \Bigl(f(X_i) - f(X_j)\Bigr) K_r(X_i,X_j)
\end{equation*}
We write $(\lambda_1(G_{n,r}),v_1(G_{n,r})),\ldots,(\lambda_n(G_{n,r}), v_n(G_{n,r}))$ for the $n$ eigenvalue/eigenvector pairs of $\Lap_{n,r}$, and adopt the usual convention of arranging the eigenvalues in ascending order, meaning $0 = \lambda_1(G_{n,r}) \leq \lambda_1(G_{n,r}) \leq \cdots \leq \lambda_1(G_{n,r})$. The Laplacian eigenmaps estimator is a nonparametric estimator which projects the data onto the span of the first $\kappa$ eigenvectors
\begin{equation*}
\wh{f}_{\LE} = \sum_{k = 1}^{\kappa} \dotp{Y}{v_k(G_{n,r})}_n v_k(G_{n,r})
\end{equation*}
where $1 \leq \kappa \leq n$ is tuning parameter. We would like to show that $f_0 \in C^1(\mc{M},Q)$ for a value of $Q$ fixed in $n$, then for an appropriate choice of $\kappa$
\begin{equation*}
\norm{\wh{f}_{\LE} - f_0}_n^2 \lesssim n^{-2/(2+m)}
\end{equation*}
with high probability.

\section{Theory}
Recall that the squared bias of $\wh{f}_{\LE}$ is upper bounded
\begin{equation}
\label{eqn:squared_bias}
\biggl\|\Ebb\Bigl[\wh{f}_{\LE} \Big | {\bf X}\Bigr] - f_0\biggr\|_n^2 \leq \frac{f_0^T \Lap_{n,r} f_0}{n \lambda_{\kappa}(G_{n,r})}
\end{equation}
(We have already supplied a sufficient bound on the variance which is independent of the dimension of $\mc{M}$). Under certain assumptions, we shall prove upper and lower bounds on the graph Sobolev seminorm $f_0^T \Lap_{n,r} f_0$ and the graph Laplacian eigenvalue $\lambda_{\kappa}(G_{n,r})$, respectively. These assumptions are as follows:
\begin{enumerate}[label=(A\arabic*)]
	\item 
	\label{asmp:max_radius}
	The radius $r$ satisfies the bounds
	\begin{equation*}
	(m + 5)d_{\infty}(P,P_n)\leq r \leq \min\biggl\{1,\frac{i_0}{10},\frac{1}{\sqrt{mK}},\frac{R}{\sqrt{27m}}\biggr\}
	\end{equation*}
	Here, $R$ is an upper bound on the reach of $\mc{M}$, $\mc{K}$ is an upper bound on the absolute values of the sectional curvatures of $\mc{M}$, $i_0$ is the injectivity radius of $\mc{M}$. 
	\item
	\label{asmp:kernel_shape}
	The kernel function $K: [0,\infty) \to [0,\infty)$ is supported on $[0,1]$ and is $Q_K$-Lipschitz continuous on its support, and satisfies $K(0) = 1$.
	\item
	\label{asmp:density}
	The density function $p$ is a $Q_p$-Lipschitz continuous function on $\mc{M}$, and is bounded away from $0$ and $\infty$,
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty
	\end{equation*}
	for all $x \in \mc{M}$.
\end{enumerate}

The bounds we prove are as follows.
\begin{enumerate}[(I)]
	\item \textbf{Graph Sobolev semi-norm:} Assume~\ref{asmp:max_radius}-\ref{asmp:density}. There exists a constant $C$ such that the following statement holds: for any $f_0 \in C^1(\mc{M},Q)$ and any $r > 0$,
	\begin{equation}
	\label{eqn:graph_sobolev_seminorm}
	f_0^T \Lap_{n,r} f_0 \leq \frac{C}{\delta}  Q^2 n^2 r^{m + 2}
	\end{equation}
	with probability at least $1 - \delta$.
	\item \textbf{Laplacian eigenvalue:}
	Assume~\ref{asmp:max_radius}-\ref{asmp:density}. There exist constants $c,C,k_{\star}$ and $\beta$ such that the following statement holds: for all $\kappa \in \mathbb{N}$ satisfying $\kappa \geq k_{\star}$, and any $r$ satisfying
	\begin{equation}
	\label{eqn:graph_eigenvalue_1}
	C \frac{\log(n)^{p_d}}{n^{1/d}} < r < c \kappa^{-1/m}
	\end{equation}
	the graph Laplacian eigenvalue satisfies the lower bound
	\begin{equation}
	\label{eqn:graph_eigenvalue_2}
	\lambda_{\kappa}(G_{n,r}) \geq c n r^{m + 2} \kappa^{2/m}.
	\end{equation}
	with probability at least $1 - C n^{-\beta}$.
\end{enumerate}

\subsection{(I): Graph Sobolev semi-norm.}

We shall prove the following upper bound on the expectation of the graph Sobolev semi-norm. 
\begin{lemma}
	\label{lem:expected_graph_sobolev_seminorm_manifold}
	Suppose assumptions~\ref{asmp:max_radius}-\ref{asmp:density} are satisfied. Then there exists a constant $C_1 > 0$ such that for any $f \in C^1(\mc{M}, Q)$,
	\begin{equation}
	\label{eqn:expected_graph_sobolev_seminorm_manifold}
	\Ebb\Bigl[f_0^T \Lap_{n,r} f_0 \Bigr] \leq C_1 p_{\max} Q^2  n^2 r^{m + 2} 
	\end{equation}
	with probability at least $1 - n\exp(c_0 n r^m p_{\min})$. 
\end{lemma}

Note that \eqref{eqn:graph_sobolev_seminorm} then follows by Markov's inequality.

\subsection{(II): Graph Laplacian eigenvalue}
Let $\Delta_{p}$ be defined for smooth functions $f: \mc{M} \to \Reals$ as
\begin{equation*}
\Delta_pf = -\frac{1}{p} \mathrm{div}\bigl(\nabla(p^2 f)\bigr)
\end{equation*}
$\Delta_p$ has a point spectrum, and we denote its eigenvalues by $\lambda_1(\mc{M}) \leq \lambda_2(\mc{M}) \leq \cdots$.

In Lemma~\ref{lem:graph_eigenvalue_manifold}, we will prove a lower bound on the graph Laplacian eigenvalues $\lambda_k(G_{n,r})$ by the eigenvalues $\lambda_k(\mc{M})$. The lower bound will be a simple corollary of the developments in \citep{trillos2019}, which we summarize in~\ref{subsec:results_of_others}.

\begin{lemma}[Lower bound on graph eigenvalue $\lambda_k(G_{n,r})$]
	\label{lem:graph_eigenvalue_manifold}
	Assume~\ref{asmp:max_radius}-\ref{asmp:density}. Then there exists some $k_{\star} \in \mathbb{N}$ such that for any $k \geq k_{\star}$, the following statement holds: if
	\begin{equation*}
	8 C_3 A \biggl(\frac{\log(n)^{p_d}}{n^{1/d}}\biggr) < r < \min\biggl\{\frac{C_2}{C_5 k^{1/m}}, \frac{1}{8C_3 L(p)}, \frac{1}{8C_3k^{1/m}}, \frac{1}{8c_3\sqrt{\mc{K}}}\biggr\}
	\end{equation*}
	then
	\begin{equation*}
	\lambda_k(G_{n,r}) \geq \frac{c_1}{2} n r^{m + 2} k^{2/m}
	\end{equation*}
	with probability at least $1 - C_4 n^{-\beta}$.
\end{lemma}

\subsection{(III): Higher order graph Sobolev semi-norms}

We have the following bound on $f^T \Lap_{n,r}^2 f$ for functions $f \in C^2(\mc{M})$.
\begin{lemma}
	\label{lem:graph_sobolev_seminorm_2}
	Assume~\ref{asmp:max_radius}-\ref{asmp:density}. Let $f \in C^2(\mc{M},Q)$. Then there exists constants $C_7$ and $c_2$ which depend only on $m,\mc{K},R$ and $i_0$ such that
	\begin{equation*}
	f^T \Lap_{n,r}^2 f \leq 2 C_7^2 Q^2 n^3 r^{2(m + 2)} p_{\max}^2
	\end{equation*}
	with probability at least $1 - 2n\exp(-2cnr^{m + 2})$. 
\end{lemma}

A similar bound holds on $f^T \Lap_{n,r}^3 f$ for functions $f \in C^3(\mc{M})$.
\begin{lemma}
	\label{lem:graph_sobolev_seminorm_3}
	Assume~\ref{asmp:max_radius}-\ref{asmp:density}. Let $f \in C^3(\mc{M},Q)$ and $p \in C^2(\mc{M},Q)$. Then there exist constants $C_8$ and $c_3$ which depend only on $m,\mc{K},R,i_0$ and $Q_K$ such that
	\begin{equation}
	\label{eqn:graph_sobolev_seminorm_3}
	f^T \Lap_{n,r}^3 f \leq \frac{C_8}{\delta} Q^2 n^4 r^{3(m + 2)} p_{\max}^3 (p_{\max} + 1)
	\end{equation}
	with probability at least $1 - 2n\exp(-c_2nr^{4 + m}) - n\exp(4 n\nu_mr^mp_{\min}/3) - \delta$. 
\end{lemma}

When we turn to $f^T \Lap_{n,r}^s f$ for $s \geq 4$, things become difficult, and at present I cannot obtain results analogous to~\ref{lem:graph_sobolev_seminorm_2} and~\ref{lem:graph_sobolev_seminorm_3}. I detail what I find difficult---in the $s = 4$ case---in Subsection~\ref{subsec:order4_is_difficult}.

Finally, we note the following bound on the graph Sobolev seminorm when $f \in H^1(P,Q)$, a direct corollary of Lemma 6 in \citep{trillos2019} (c.f. Lemma 3.3 \citep{burago2014}) plus Markov's inequality.

\begin{lemma}
	Assume~\ref{asmp:max_radius}-\ref{asmp:density}. Then there exists a universal constant $C > 0$ such that the following statement holds for any $f \in H^1(P)$:
	\begin{equation*}
	f^T \Lap_{n,r} f \leq \frac{1}{\delta}\bigl(1 + Q_p r\bigr) \bigl(1 + CmKr^2\bigr) \sigma_K r^{m + 2} |f|_{H^1(P)}
	\end{equation*}
	with probability at least $1 - \delta$. 
\end{lemma}

\section{Proof of Theorems and Major Lemmas}
We will often make use of the following geometric estimates, which hold under assumption~\ref{asmp:max_radius}. These help us to convert from integrals over $\mc{M}$ to integrals over a Euclidean space. First
\begin{equation}
\label{eqn:geodisic_euclidean_distance}
d_{\mc{M}}(y,x) \leq \norm{x - y}_{\Rd} + \frac{8}{R^2} \norm{x - y}_{\Rd}^3
\end{equation} 
and additionally
\begin{equation}
\label{eqn:manifold_ball_volume}
\abs{\mu(B(x,r)) - \nu_m r^m} \leq C_0r^{m + 2}
\end{equation}
See ((3.3) and (3.2) of (Calder and Garcia-Trillos 19)).

\subsection{Proof of Lemma~\ref{lem:expected_graph_sobolev_seminorm_manifold}}
In the following manipulations we invoke first the Holder property of $f$, then the upper bound in~\ref{eqn:geodisic_euclidean_distance}, and finally Lemma~\ref{lem:max_degree} with $\delta = 2$ to get
\begin{align*}
f^T \Lap_{n,r}f  & = \sum_{i,j = 1}^{n} \Bigl(f(X_i) - f(X_j)\Bigr)^2 K\biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\biggr) \\
& \leq Q^2 \sum_{i,j = 1}^{n} \Bigl(d_{\mc{M}}(X_i,X_j)\Bigr)^2 K\biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\biggr) \\
& \leq 2 Q^2 \sum_{i,j = 1}^{n} \Bigl(\norm{X_i - X_j}_{\Rd}^2 + \frac{64}{R^4} \norm{X_i - X_j}_{\Rd}^6\Bigr) K\biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\biggr) \\ 
& \leq 2 Q^2 \Bigl(r^2 + \frac{64}{R^4} r^6\Bigr) n \deg_{\max}(G_{n,r}) \\
& \leq 4 Q^2 p_{\max}\Bigl(r^2 + \frac{64}{R^4} r^6\Bigr) \Bigl(\nu_mr^m + C_0r^{m + 2}\Bigr) n^2  
\end{align*}
with the last inequality holding with probability at least $1 - n\exp(-\frac{4}{3}n\nu_mr^mp_{\min})$. Lemma~\ref{lem:expected_graph_sobolev_seminorm_manifold} then follows from appropriate choice of constants $c_0$ and $C_1$.
\subsection{Proof of Lemma~\ref{lem:graph_eigenvalue_manifold}}

Applying Theorem~\ref{thm:trillos19_2}---which gives an estimate on the $\infty$-optimal transport distance between $P$ and $P_n$---to Theorem~\ref{thm:trillos19_1}---which gives a lower bound on $\lambda_k(G_{n,r})$ in terms of $\lambda_k(\mc{M})$ and $d_{\infty}(P,P_n)$--- we have that if
\begin{equation*}
2C_3 A \frac{\log(n)^{p_d}}{n^{1/d}} \leq r < \min\biggl\{\frac{C_2}{\sqrt{\lambda_k(\mc{M})}}, \frac{1}{8C_3 L(p)}, \frac{1}{8C_3\sqrt{\lambda_k(\mc{M})}}, \frac{1}{8C_3\sqrt{\mc{K}}}\biggr\}
\end{equation*}
then
\begin{equation*}
\lambda_k(G_{n,r}) \geq \frac{1}{2} nr^{m+2}\lambda_k(\mc{M})
\end{equation*}
with probability at least $1 - C_4 n^{-\beta}$. Lemma~\ref{lem:graph_eigenvalue_manifold} follows from replacing all instances of $\lambda_k(\mc{M})$ in the above expression with estimates given by Weyl's Law (Theorem~\ref{thm:weyl}).

\subsection{Proof of Lemma~\ref{lem:graph_sobolev_seminorm_2}}
Note that
\begin{equation*}
f^T \Lap_{n,r}^2 f = \sum_{i = 1}^{n} \Bigl(\Lap_{n,r}f(X_i)\Bigr)^2
\end{equation*}
Lemma~\ref{lem:graph_sobolev_seminorm_2} follows directly from the pointwise estimate Lemma~\ref{lem:pointwise_laplacian_ub} upon taking a union bound over $x = X_1,\ldots,X_n$.

\subsection{Proof of Lemma~\ref{lem:graph_sobolev_seminorm_3}}
We will consider $\Lap_{n,r}f(X_i)$ as an estimate of $\sigma_K \Delta_pf(X_i)$ (when appropriately scaled), and show that this deviation is small when $r$ is sufficiently large.  To begin, writing $\Lap_{n,r}f(X_i) = \sigma_{K}\Delta_{p}f(X_i)nr^{m + 2} + \Lap_{n,r}f(X_i) - \sigma_{K}\Delta_{p}f(X_i)nr^{m+2}$ inside the definition of the 3rd-order graph Sobolev seminorm, we obtain
\begin{align*}
f^T \Lap_{n,r}^3 f & = \sum_{i,j = 1}^{n} \Bigl(\Lap_{n,r}f(X_i) - \Lap_{n,r}f(X_j)\Bigr)^2 K_r(X_i,X_j) \\
& \leq 6 \sum_{i,j = 1}^{n} \Bigl(\Lap_{n,r}f(X_i) - \sigma_{K}\Delta_{p}f(X_i)nr^{m+2}\Bigr)^2 K_r(X_i,X_j)~+ \\
&~~3 \sigma_K^2 n^2 r^{2(m + 2)} \sum_{i,j = 1}^{n} \Bigl(\Delta_pf(X_i) - \Delta_pf(X_j)\Bigr)^2 K_r(X_i,X_j)
\end{align*}
Applying Theorem~\ref{thm:calder19_1} with $\delta = r$, and Lemma~\ref{lem:max_degree} with $\delta = 1$, we have
\begin{align*}
\sum_{i,j = 1}^{n} \Bigl(\Lap_{n,r}f(X_i) - \sigma_{K}\Delta_{p}f(X_i)nr^{m+2}\Bigr)^2  K_r(X_i,X_j) & \leq  C_6^2 \bigl([f]_{1;\mc{M}} + \norm{f}_{C^3(\mc{M})} + 1\bigr)^2 n^2r^{2m + 6} \sum_{i,j = 1}^{n} K_r(X_i,X_j) \\
& \leq  C_6^2 \bigl([f]_{1;\mc{M}} + \norm{f}_{C^3(\mc{M})} + \norm{p}_{C^2(\mc{M})}\bigr)^2 n^3r^{2m + 6} \deg_{\max}(G_{n,r}) \\
& \leq 4 C_6^2 \bigl([f]_{1;\mc{M}} + \norm{f}_{C^3(\mc{M})} + \norm{p}_{C^2(\mc{M})}\bigr)^2 n^4 r^{2m + 6} \Bigl[\nu_mr^m + C_0r^{m + 2}\Bigr] p_{\max} \\
& \leq C p_{\max}\bigl(Q + p_{\max}\bigr)^2n^4 r^{3m + 6},
\end{align*}
with probability at least $1 - 2n\exp(-c_2nr^{4 + m}) - n\exp(4 n\nu_mr^mp_{\min}/3)$.

It remains to upper bound $(\Delta_pf)^T \Lap_{n,r} (\Delta_pf)$. Rewriting
\begin{equation*}
\Delta_pf = -\bigl(\nabla p^T \nabla f + p \Delta_{\mc{M}} f\bigr)
\end{equation*}
we observe that since $f \in C^3(\mc{M}, Q)$ and $p \in C^2(\mc{M}, p_{\max})$, $\Delta_pf \in C^1(\mc{M},4Qp_{\max})$. By Lemma~\ref{lem:expected_graph_sobolev_seminorm_manifold} we therefore have
\begin{equation*}
(\Delta_pf)^T \Lap_{n,r} (\Delta_pf) \leq \frac{16}{\delta} C_1 n^2 r^{m + 2} Q^2 p_{\max}^4
\end{equation*}
with probability at least $1 - \delta$. The claim of Lemma~\ref{lem:graph_sobolev_seminorm_3} then follows after an appropriate choice of constant in~\eqref{eqn:graph_sobolev_seminorm_3}.

\section{Additional Theory}

\subsection{Estimates on Graph Eigenvalues}
\label{subsec:results_of_others}

\begin{theorem}[(Part of) Theorem 4 of \citep{trillos2019}]
	\label{thm:trillos19_1}
	There exist constants $C_2$ and $C_3$ which depend only on $m, p_{\min}, p_{\max}, L(p)$ and $K$ such that if 
	\begin{equation*}
	\sqrt{\lambda_k(\mc{M})} r < C_2,
	\end{equation*}
	then,
	\begin{equation*}
	\lambda_k(G_{n,r}) \geq n r^{m + 2}\lambda_k(\mc{M}) \Biggl[1 - C_3\biggl(L(p)r + \frac{d_{\infty}(P,P_n)}{r} + \sqrt{\lambda_k(\mc{M})}r + \mc{K}r^2\biggr)\Biggr]
	\end{equation*}
\end{theorem}

In order to make use of Theorem~\ref{thm:trillos19_1}, we need an upper bound on $d_{\infty}(P,P_n)$. Theorem~\ref{thm:trillos19_2} provides a probabilistic estimate.

\begin{theorem}[Theorem 2 of \citep{trillos2019}]
	\label{thm:trillos19_2}
	Let $P$ satisfy~\ref{asmp:density}. Then for any $\beta > 1$ and every $n \in \mathbb{N}$, there exists a transport map $T_n: \mc{M} \to {\bf X}$ and a constant $A$ such that
	\begin{equation*}
	\sup_{x \in \mc{M}} d_{\infty}(x,T_n(x)) \leq A \cdot 
	\begin{cases*}
	\frac{(\log(n))^{3/4}}{n^{1/2}},& ~~\textrm{if $m = 2$,} \\
	\biggl(\frac{\log(n)}{n}\biggr)^{1/m},& ~~\textrm{if $m \geq 3$,}
	\end{cases*}
	\end{equation*}
	with probability at least $1 - C_4 n^{-\beta}$. The constant $A$ depends only on $\mc{K},i_0,m,\mu(\mc{M}), \alpha$ and $\beta$, and the constant $C_4$ depends only on $\mc{K},i_0,m,\mu(\mc{M})$.
\end{theorem}

\subsection{Weyl's Law}

Although Weyl's Law is traditionally stated with respect to the unweighted Laplace-Beltrami operator, the same asymptotics apply to $\Delta_{p}$ when $p$ is bounded away from $0$ and $\infty$ on $\mc{M}$. Let $N(\lambda)$ count the number of eigenvalues of $\mc{M}$ which are less than $\lambda$.
\begin{theorem}[Weyl's Law]
	\label{thm:weyl}
	Assume~\ref{asmp:density}. If $\mc{M}$ is a compact connected oriented manifold then
	\begin{equation*}
	N(\lambda) \sim \lambda^{m/2}
	\end{equation*}
	or, equivalently,
	\begin{equation*}
	\lambda_k(\mc{M}) \sim k^{2/m}.
	\end{equation*}
\end{theorem}

\subsection{Pointwise estimates on graph Laplacians}

We give some pointwise estimates on $\Lap_{n,r}f(x)$, many of which are reproduced from \citep{calder2019}. Define
\begin{equation*}
\Lap_{P,r}f(x) := \frac{1}{r^{m + 2}} \int_{B(x,r)} K_r(y,x) \Bigl( f(x) - f(y) \Bigr) p(y) \,d\mu(y)
\end{equation*}
to be a nonlocal Laplacian operator, which acts as an intermediary between $\Lap_{n,r}$ and $\Delta_p$.

\begin{theorem}[Theorem 3.3 of \citep{calder2019}]
	\label{thm:calder19_1}
	Assume~\ref{asmp:max_radius}-\ref{asmp:density}. Let $f \in C^3(\mc{M})$ and $p \in C^2(\mc{M})$. There exist constants \textcolor{red}{$C_6$} and $c_2$ such that for any $r \leq \delta \leq r^{-1}$, 
	\begin{equation*}
	\max_{1 \leq i \leq n} \abs{\Lap_{n,r}f(X_i) - \sigma_{K} \Delta_{p} f(X_i) nr^{m + 2}} \leq C_6 \bigl([f]_{1;\mc{M}} + \norm{f}_{C^3(\mc{M})} + \norm{p}_{C^2(\mc{M})}\bigr) \delta  nr^{m + 2} 
	\end{equation*}
	with probability at least $1 - 2n\exp(-c\delta^2 nr^{m+2})$.
\end{theorem}

\begin{lemma}[Lemma 3.4 of \citep{calder2019}]
	\label{lem:calder19}
	Assume~\ref{asmp:max_radius}-\ref{asmp:density}. Let $f \in C^1(\mc{M})$. Then for $x \in \mc{M}$ and $r \leq \delta \leq r^{-1}$,
	there exists constants $C_6$ and $c_2$ which depend only on $m,\mc{K},R$ and $i_0$ such that
	\begin{equation*}
	\Pbb\biggl[\Bigl|\frac{1}{nr^{m + 2}} \Lap_{n,r}f(x) - \Lap_{P,r}f(x)\Bigr| \geq C_6 [f]_{1,B(x,2r)} \delta \biggr] \leq 2\exp(-c_2\delta^2 n r^{m + 2})
	\end{equation*}
\end{lemma}

Most analysis on $\Lap_{n,r}f(x)$ provide pointwise estimates assuming $f \in C^{q + 2}(\mc{M})$ for some $q > 0$. We shall instead assume only $f \in C^2(\mc{M})$, and content ourselves with merely an upper bound on $|\Lap_{n,r}f(x)|$.
\begin{lemma}
	\label{lem:pointwise_laplacian_ub}
	Assume~\ref{asmp:max_radius}-\ref{asmp:density}. Let $f \in C^2(\mc{M})$. Then for $x \in \mc{M}$ and $r \leq \delta \leq r^{-1}$, there exists constants $C_7$ and $c_2$ which depend only on $m,\mc{K},R$ and $i_0$ such that
	\begin{equation*}
	\Bigl|\Lap_{n,r}f(x)\Bigr| \leq C_7 nr^{m + 2}\Bigl([f]_{1;B(x,2r)}\delta + p_{\max} \nu_d[f]_{1,B(x,r)} r + \norm{\nabla f(x)} +  \norm{f}_{C^2(B(x,r))}\Bigr)
	\end{equation*}
	with probability at least $1 - 2\exp(-c_2\delta^2 n r^{m + 2})$.
\end{lemma}
\begin{proof}
	We shall follow the proof of Theorem 3.3 of~\citep{calder2019}, deviating when necessary to fit our differing assumptions and ultimate goal. Let
	\begin{equation*}
	\Lap_{P,r}^{(i)}f(x) = \frac{1}{r^{m + 2}} \int_{B(x,r)} K\biggl(\frac{d_{\mc{M}}(y,x)}{r}\biggr)\bigl(f(x) - f(y)\bigr) p(y) \,d\mu(y)
	\end{equation*} 
	be an intrinsic analogue to $\Lap_{P,r}$. It follows from~\eqref{eqn:geodisic_euclidean_distance} and~\ref{asmp:kernel_shape} that 
	\begin{align}
	\abs{\Lap_{P,r}^{(i)}f(x) - \Lap_{P,r}f(x)} & \leq \frac{1}{r^{2 + m}} \int_{B(x,r)} \Biggl|K(\frac{d_{\mc{M}}(x,y)}{r}) - K(\frac{d_{\mc{M}}(x,y)}{r})\Biggr| \bigl|f(x) - f(y)\bigr| p(y) \,d\mu(y) \nonumber \\ 
	& \leq \frac{p_{\max}}{r^{m}} \int_{B(x,r)} \bigl|f(x) - f(y)\bigr| \,d\mu(y)\\
	& \leq p_{\max} \nu_d[f]_{1,B(x,r)} r \label{eqn:pointwise_laplacian_ub}
	\end{align}
	By assumption $r < i_0$, and therefore the exponential map $\exp_x: B(0,r) \subset \mc{T}_{x}(\mc{M}) \to \mc{Mc}$ is a diffeomorphism. For $v \in B(0,r)$ let $w(v) = f(\exp_x(v))$ and $\rho(v) = p(\exp_x(v))$, i.e. express $f$ and $p$ in terms of normal Riemmanian coordinates,  and let $J_x(v)$ be the Jacobian of $\exp_x$ at $v$.  Then
	\begin{align*}
	\Lap_{P,r}^{(i)}f(x) & = -\frac{1}{r^{m + 2}} \int_{B(0,r)} K\biggl(\frac{\norm{v}}{r}\biggr)\bigl(w(v) - w(0)\bigr) \rho(v) J_x(v) \,d\mu(v) \\
	& = -\frac{1}{r^2} \int_{B(0,1)} K(\norm{v})  \bigl(w(rv) - w(0)) \rho(rv) J_x(rv) \,d\mu(v).
	\end{align*}
	Now we plug in the Taylor expansions $w(rv) = w(0) + \nabla w(0) \cdot v + O(\norm{w}_{C^2(B(0,r))}r^2)$, $\rho(rv) = \rho(0) + O(r)$ and $J_x(rv) = 1 + O(r^2)$ to obtain
	\begin{align}
	\Lap_{P,r}^{(i)}f(x) & = -\frac{1}{r^2} \int_{B(0,1)} K(\norm{v})  \Bigl(r \nabla w(0)^T v + O\bigl(\norm{w}_{C^2(B(0,r))}r^2\bigr)\Bigr) \Bigl(\rho(0) + O(r)\Bigr) \Bigl(1 + O(r^2)\Bigr) \,d\mu(v) \nonumber \\
	& = O\Bigl(\norm{\nabla w(0)} +  \norm{w}_{C^2(B(0,r))} \Bigr) \label{eqn:pointwise_laplacian_ub_2}
	\end{align}
	The claim then follows by combining~\eqref{eqn:pointwise_laplacian_ub},~\eqref{eqn:pointwise_laplacian_ub_2} and Lemma~\ref{lem:calder19}.
\end{proof}

\subsection{Degree bounds}

Lemma~\ref{lem:max_degree} follows from the multiplicative form of Hoeffding's inequality along with~\eqref{eqn:manifold_ball_volume} and~\ref{asmp:kernel_shape}.
\begin{lemma}
	\label{lem:max_degree}
	Assume~\ref{asmp:max_radius}-\ref{asmp:density}. Then
	\begin{equation*}
	\deg_{\max}(G_{n,r}) \leq (1 + \delta)\Bigl[\nu_mr^m + C_0r^{m + 2}\Bigr] p_{\max} n
	\end{equation*}
	with probability at least $1 - n\exp(-\delta^2 n \nu_m r^m p_{\min}/3)$.
\end{lemma}

\subsection{Explaining what is difficult when $s = 4$.}
\label{subsec:order4_is_difficult}
Let me start by giving a rough summary of the difficulty, and then go into more detail .We would like to show
\begin{equation}
\label{eqn:order4_is_difficult_1}
f^T \Lap_{n,r}^4 f = \sum_{i = 1}^{n} \Bigl(\Lap_{n,r}^2f(X_i)\Bigr)^2 \lesssim n^{5}r^{8(m + 2)}
\end{equation}
for $f \in C^4(\mc{M})$.
Note that
\begin{equation*}
\Lap_{P,r}f(x) = \int_{B(x,r) \cap \mc{M}} \bigl(f(x) - f(y)\bigr) \,dP(y)
\end{equation*}
is (up to scaling by a factor of $nr^{m+2}$)the expectation of $\Lap_{n,r}f(x)$ . It is possible to show a sufficiently tight bound on $|\Lap_{n,r}^2f(x) - nr^{2+m}\Lap_{P,r}^2f(x)|$ when $r \gg n^{-1/(6 + m)}$, thus handling the variance. To handle the bias, we can show that
\begin{equation*}
\Lap_{P,r}f(x) = nr^{m + 2} \sigma_{K} \Delta_pf(x) + \mc{I}f(x)
\end{equation*}
where $\Delta_p$ is the weighted Laplace-Beltrami operator on $\mc{M}$, and $\mc{I}f$ is an error term satisfying $\mc{I}f(x) = O(r^{m + 3})$. This bound on the error term is sufficient for $s = 1$ up to $s = 3$, however, when $s \geq 4$ it is insufficient (check by plugging back in to~\eqref{eqn:order4_is_difficult_1}). In the Euclidean case, we had that $\mc{I}f(x)$ itself belong to $C^1$ and thus could argue $(\Lap_{P,r} \mc{I}f)(x) \ll \mc{I}f(x)$. In the manifold case, I no longer know how to do this.

The longer version: we can make the following progress towards upper bound $f^T \Lap_{n,r}^4 f$:
\begin{align*}
f^T \Lap_{n,r}^4 f & \leq n \max_{1 \leq i \leq n} \abs{\Lap_{n,r}^2 f(X_i)}^2 \\
& \leq n \max_{1 \leq i \leq n} \abs{\Lap_{n,r}(\Lap_{n,r} - nr^{m + 2}\Lap_{P,r})f(X_i)}^2 + n^2 r^{m + 2} \max_{1 \leq i \leq n} \abs{(\Lap_{n,r} - nr^{m + 2}\Lap_{P,r})\Lap_{P,r}f(X_i)}^2 +~\\
& ~~n^4 r^{4m + 8} \max_{1 \leq i \leq n} \abs{\Lap_{P,r}^2f(X_i)}
\end{align*}
We can upper bound the first term as follows:
\begin{align*}
\abs{\Lap_{n,r}(\Lap_{n,r} - nr^{m + 2}\Lap_{P,r})f(x)}^2 & \leq \max_{1 \leq j \leq n} \abs{(\Lap_{n,r} - nr^{m + 2}\Lap_{P,r})f(X_j)} \cdot \sum_{j = 1}^{n} K\Bigl(\frac{\norm{x - X_j}_{\Rd}}{r}\Bigr) \\
& \leq C nr^m \max_{1 \leq j \leq n} \abs{(\Lap_{n,r} - nr^{m + 2}\Lap_{P,r})f(X_j)} \\
& \leq C n^2r^{2m + 2} \delta 
\end{align*}
where the last inequality follows from Bernstein's inequality and holds with probability at least on the order of $1 - \exp(-nr^{2 + m}\delta^2)$. Choosing $\delta \asymp r^2$ gives the desired rate, assuming $n \gg r^{6 + m}$.

We can use a similar argument to upper bound the second term.

The third term---$\Lap_{P,r}^2f(X_i)$---is what poses the challenge. We want to convert the integral on $\mc{M}$ to an integral on a Euclidean space using exponential maps. Letting $\exp_x:T_x(\mc{M}) \to \mc{M}$ be the exponential map, letting $\wt{B}$ satisfy $\exp_x(\wt{B}) = B(x,r) \cap \mc{M}$, associating the tangent plane $T_x(\mc{M})$ with $\Reals^m$, and writing $g_x(v) = \norm{x - \exp_x(v)}_{\Rd} - \norm{v}_{\Reals^m}$ we have
\begin{align*}
\Lap_{P,r}f(x) & = \int_{\mc{M}} (f(x) - f(y)) K\Bigl(\frac{\norm{y - x}_{\Rd}}{r}\Bigr) p(y) \,d\mu(y) \\
& = \int_{\wt{B}} (w(0) - w(v)) K\Bigl(\frac{\norm{x - \exp_x(v)}_{\Rd}}{r}\Bigr) \rho(v) J_v(x) \,dv \\
& = \int_{\wt{B}} (w(0) - w(v)) K\Bigl(\frac{\norm{v}_{\Reals^m} + g_x(v)}{r}\Bigr) \rho(v) J_v(x) \,dv \\
& = \int_{B(0,r) \subset \Reals^m} (w(0) - w(v)) K\Bigl(\frac{\norm{v}_{\Reals^m}}{r}\Bigr) \rho(v) J_v(x) \,dv +~~\\
&~~ \underbrace{\int_{\wt{B}} (w(0) - w(v)) \biggl[K\Bigl(\frac{\norm{v}_{\Reals^m} + g_x(v)}{r}\Bigr) - K\Bigl(\frac{\norm{v}_{\Reals^m}}{r}\Bigr)\biggr] \rho(v) J_v(x) \,dv}_{\mc{I}(x)}
\end{align*}
The first of the two integrals is now an integral over a ball in a Euclidean domain, and is not a problem. The second integral is tricky. The only way I know how to analyze $\mc{I}(x)$ is to use the following facts: first, $\abs{w(v) - w(0)} \lesssim r$, second that $g_x(v) \lesssim r^3$ for all $\exp_x(v) \in B(0,r) \cap \mc{M}$, third that $\abs{\mu(B(x,r) \cap {\mc{M}}) - \nu_m r^m} \lesssim r^{m + 2}$, and fourth that $\abs{J_x(v)} \lesssim 1 + r^2$. Together, these give
\begin{align*}
\int_{\wt{B}} (w(0) - w(v)) \biggl[K\Bigl(\frac{\norm{v}_{\Reals^m} + g_x(v)}{r}\Bigr) - K\Bigl(\frac{\norm{v}_{\Reals^m}}{r}\Bigr)\biggr] \rho(v) J_v(x) \,dv  & \lesssim r \Bigl(\mu\bigl(B(x,r + r^3) \cap {\mc{M}} \bigr) - \mu\bigl(B(x,r) \cap {\mc{M}}\bigr)\Bigr)  \\
& \lesssim r^{m + 3}.
\end{align*}
But this only gives $\Lap_{P,r}^2f(x) \lesssim r^{2m + 3}$, which is not nearly enough. The problem is that I am not using any smoothness properties of $\mc{I}(x)$ in my upper bound on $\Lap_{P,r}\mc{I}f(x)$. I suspect it may have such properties, related to the smoothness assumptions we place on $\mc{M}$, but proving it seems non-trivial.

\section{Notation}

\begin{itemize}
	\item For any $x,y \in \mc{M}$ we write $d_{\mc{M}}(x,y)$ for the geodesic distance between $x$ and $y$, that is the length of the shortest path connecting $x$ and $y$. 
	\item For measures $P$ and $Q$ on $\mc{M}$, we write $d_{\infty}(P,Q)$ for the $\infty$-optimal transport distance between $P$ and $Q$. Formally,
	\begin{equation*}
	d_{\infty}(P,Q) = \inf\biggl\{ \mathrm{esssup}_{\gamma}\Bigl\{\abs{x - y}: x,y \in \mc{M} \Bigr\} : \gamma \in \Gamma(P,Q) \biggr\}
	\end{equation*}
	where $\Gamma(P,Q)$ is the set of joint distributions on $\mc{M} \times \mc{M}$ with marginals $P$ and $Q$.

\end{itemize}

\bibliographystyle{plainnat}
\bibliography{../../graph_testing_bibliography}

\end{document}

### 2020-08-19 notes/weekly_notes/4.2.20.NotesForWeek.tex
Added weekly notes, which include some derivations of testing results for the fixed graph case.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\makeatletter
\newcommand{\vast}{\bBigg@{4}}
\newcommand{\Vast}{\bBigg@{5}}
\makeatother

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}
\newcommand{\Lap}{{\bf L}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{\mathcal{L}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LS}{\mathrm{LS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 3/27/20 - 4/2/20}
\author{Alden Green}
\date{\today}
\maketitle

We observe data $X_1,\ldots,X_n \sim P$, where $P$ is defined over a sample space $\Xset \subset \Reals^d$. For a given kernel $K: \Reals^+ \to \Reals^+$, and a radius $r$, define $K_r(z) = K(z/r)$. Construct the neighborhood graph $G_{n,r} = ([n],W)$, where $W_{ij} = K_r(\norm{X_i - X_j}_2)$. The neighborhood graph Laplacian $\Lap_{n,r}$ is then defined to be $D - W$, where $D$ is the diagonal degree matrix of $G_{n,r}$ with entries $D_{ii} = \sum_{j = 1}^{n} W_{ij}$.

Let $f: \Xset \to \Reals$ possess $s$ derivatives, which we will eventually assume are well-behaved in some sense (e.g. $f$ is an the order-$s$ Holder or Sobolev class). Our goal is to understand the behavior of three statistics: first, the pointwise evaluation $\Lap_{n,r}^sf(x)$ for some $x \in \Xset$ (which we will formally define momentarily); second the pointwise evaluation $\Lap_{n,r}^sf(X_i)$ for some $i = 1,\ldots,n$; and third the seminorm $f^T \Lap_{n,r}^s f$.

\section{Pointwise Evaluation}

For any $x \in \Xset$, define
\begin{equation}
\label{eqn:pointwise_evaluation}
\Bigl(\Lap_{n,r}f\Bigr)(x) := \sum_{i = 1}^{n} \Bigl( f(x) - f(X_i) \Bigr) K_r(x,X_i).
\end{equation}
When $x = X_i$ for $i = 1,\ldots,n$, we have $\Bigl(\Lap_{n,r}f\Bigr)(X_i) = \Bigl(\Lap_{n,r}f\Bigr)_i$, so ~\eqref{eqn:pointwise_evaluation} defines an extension of $\Lap_{n,r}$ from the data $X_1,\ldots,X_n$ to all of $\Xset$. For $s > 1$, recursively define
\begin{equation*}
\Bigl(\Lap_{n,r}^s f\Bigr)(x) := \sum_{i = 1}^{n} \Bigl\{\Bigl(\Lap_{n,r}^{s-1}f\Bigr)(x) - \Bigl(\Lap_{n,r}^{s-1}f\Bigr)(X_i)\Bigr\}K_r(x,X_i)
\end{equation*}
Our goal is to show (a) compute the expectation of $\bigl(\Lap_{n,r}^s f\bigr)(x)$, and (b) to show that $\bigl(\Lap_{n,r}^s f\bigr)(x)$ concentrates around its expectation. We will begin with the simplest non-trivial case of $s = 2$.

\subsection{$s = 2$}

The evaluation $\bigl(\Lap_{n,r}^2 f\bigr)(x)$ can be written as
\begin{equation}
\label{eqn:squared_Laplacian}
\bigl(\Lap_{n,r}^2 f\bigr)(x) = \sum_{i = 1}^{n} \sum_{j = 1}^{n} \bigl(f(x) - f(X_j)\bigr)K_r(x,X_j)K_r(x,X_i) - \sum_{i = 1}^{n} \sum_{j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)K_r(X_i,X_j)K_r(x,X_i)
\end{equation}
\subsubsection{Expectation at a fixed point.}
From~\eqref{eqn:squared_Laplacian}, we see
\begin{equation*}
\Ebb\Bigl[\bigl(\Lap_{n,r}^2 f\bigr)(x)\Bigr] = \sum_{i = 1}^{n} \sum_{j = 1}^{n} \biggl\{\Ebb\Bigl[\bigl(f(x) - f(X_j)\bigr)K_r(x,X_j)K_r(x,X_i)\Bigr] - \Ebb\Bigl[\bigl(f(X_i) - f(X_j)\bigr)K_r(X_i,X_j)K_r(x,X_i)\Bigr]\biggr\}
\end{equation*}
The expectation of the summand depends on whether $i = j$. If $i = j$, then 
\begin{equation*}
\Ebb\Bigl[\bigl(f(x) - f(X_j)\bigr)K_r(x,X_j)K_r(x,X_i)\Bigr] = \int \bigl(f(x) - f(y)\bigr) \bigl(K_r(\norm{y - x})\bigr)^2 \,dP(y) =: \bigl(I_{1,P}f\bigr)(x)
\end{equation*}
and clearly $\Ebb\Bigl[\bigl(f(X_i) - f(X_j)\bigr)K_r(X_i,X_j)K_r(x,X_i)\Bigr] = 0$. Otherwise if $i \neq j$,
\begin{equation*}
\Ebb\Bigl[\bigl(f(x) - f(X_j)\bigr)K_r(x,X_j)K_r(x,X_i)\Bigr] = \Ebb \Bigl[\bigl(f(x) - f(X)\bigr)K_r(x,X)\Bigr] \cdot \Ebb\Bigl[K_r(x,X)\Bigr] =: \bigl(L_{P,r}f\bigr)(x) \cdot \Ebb\Bigl[K_r(x,X)\Bigr],
\end{equation*}
by the law of conditional expectation
\begin{equation*}
\Ebb\Bigl[\bigl(f(X_i) - f(X_j)\bigr)K_r(X_i,X_j)K_r(x,X_i)\Bigr] = \Ebb\Bigl[\bigl(L_{P,r}f\bigr)(X) \cdot K_r(x,X)\Bigr],
\end{equation*}
and therefore
\begin{equation*}
\Ebb\Bigl[\bigl(f(x) - f(X_j)\bigr)K_r(x,X_j)K_r(x,X_i)\Bigr] -\Ebb\Bigl[\bigl(f(X_i) - f(X_j)\bigr)K_r(X_i,X_j)K_r(x,X_i)\Bigr] = \bigl(L_{P,r}^2f\bigr)(x)
\end{equation*}
We conclude that
\begin{equation*}
\Ebb\Bigl[\bigl(\Lap_{n,r}^2 f\bigr)(x)\Bigr] = n(n - 1) \bigl(L_{P,r}^2f\bigr)(x) + n \bigl(I_{1,P}f\bigr)(x)
\end{equation*}

\section{Semi-norm}

\subsection{$s = 3$}

We use the notation $\Ebb_{-i,j}[X] = \Ebb[X|X_i,X_j]$. The third-order graph Sobolev seminorm $f^T L_{n,r}^3 f$ can be written as
\begin{align*}
f^T L_{n,r}^3 f & = \sum_{i,j = 1}^{n} \Bigl(\bigl(L_{n,r}f\bigr)(X_i) - \bigl(L_{n,r}f\bigr)(X_j)\Bigr)^2 K_r\Bigl(\norm{X_i - X_j}_2\Bigr) \\
& \leq 3 \sum_{i,j = 1}^{n} \biggl(\bigl(L_{n,r}f\bigr)(X_i) - \Ebb_{-i,j}\Bigl[\bigl(L_{n,r}f\bigr)(X_i)\Bigr]\biggr)^2 K_r\Bigl(\norm{X_i - X_j}_2\Bigr) ~~ + \\
& ~~~~3 \sum_{i,j = 1}^{n} \biggl(\bigl(L_{n,r}f\bigr)(X_j) - \Ebb_{-i,j}\Bigl[\bigl(L_{n,r}f\bigr)(X_j)\Bigr]\biggr)^2 K_r\Bigl(\norm{X_i - X_j}_2\Bigr)~~ + \\
& ~~~~3 \sum_{i,j = 1}^{n} \biggl(\bigl(\Ebb_{-i,j}\Bigl[\bigl(L_{n,r}f\bigr)(X_i) \Bigr] - \Ebb_{-i,j}\Bigl[\bigl(L_{n,r}f\bigr)(X_j)\Bigr]\biggr)^2 K_r\Bigl(\norm{X_i - X_j}_2\Bigr)
\end{align*}
We upper bound the expectation of each of the three terms in the summand on the right hand side.

\textit{Term 1.}

By the law of iterated expectation,
\begin{align*}
& \Ebb \Biggl[\biggl(\bigl(L_{n,r}f\bigr)(X_i) - \Ebb_{-i,j}\Bigl[\bigl(L_{n,r}f\bigr)(X_i)\Bigr]\biggr)^2 K_r\Bigl(\norm{X_i - X_j}_2\Bigr)\Biggr] = \\
& ~~~ \Ebb \Biggl[ \Ebb_{-i,j}\biggl[ \biggl(\bigl(L_{n,r}f\bigr)(X_i) - \Ebb_{-i,j}\Bigl[\bigl(L_{n,r}f\bigr)(X_i)\Bigr]\biggr)^2 \biggr]   K_r\Bigl(\norm{X_i - X_j}_2\Bigr)\Biggr] = \\
& ~~~ \Ebb \Biggl[ \Var_{-i,j}\Bigl[\bigl(L_{n,r}f\bigr)(X_i) \Bigr]   K_r\Bigl(\norm{X_i - X_j}_2\Bigr)\Biggr]
\end{align*}
We express the conditional variance as a sum of conditional covariances,
\begin{align*}
\Var_{-i,j}\biggl[ \Bigl(\bigl(L_{n,r}f\bigr)(X_i)\Bigr) \biggr] & = \sum_{k = 1}^{n} \sum_{\ell = 1}^{n} \Cov_{-i,j}\biggl[\Bigl(f(X_i) - f(X_k)\Bigr)K_r\Bigl(\norm{X_i - X_k}_2\Bigr), \Bigl(f(X_i) - f(X_{\ell})\Bigr)K_r\Bigl(\norm{X_i - X_{\ell}}_2\Bigr)\biggr] \\
& = (n - 3) \Var_{-i}\biggl[\Bigl(f(X_i) - f(X)\Bigr)K_r\Bigl(\norm{X_i - X}_2\Bigr)\biggr].
\end{align*}
where in the second equality we have used that $X_1,\ldots,X_n$ are i.i.d samples, and so conditional on $X_i$ and $X_j$ the covariance is equal to zero unless $k = \ell$. By Lemma~\ref{lem:expectation_1},
\begin{align*}
\Var_{-i}\biggl[\Bigl(f(X_i) - f(X)\Bigr)K_r\Bigl(\norm{X_i - X}_2\Bigr)\biggr] \leq \Ebb_{-i}\Biggl[\biggl(\Bigl(f(X_i) - f(X)\Bigr)K_r\Bigl(\norm{X_i - X}_2\Bigr)\biggr)^2\Biggr] \leq M^2 r^{2+d}K_{\max}^2 p_{\max}
\end{align*}
and as a result the expectation of term 1 is upper bounded
\begin{align*}
\Ebb \Biggl[\biggl(\bigl(L_{n,r}f\bigr)(X_i) - \Ebb_{-i,j}\Bigl[\bigl(L_{n,r}f\bigr)(X_i)\Bigr]\biggr)^2 K_r\Bigl(\norm{X_i - X_j}_2\Bigr)\Biggr] & \leq M^2K_{\max}^2 p_{\max} \cdot nr^{2 + d} \Ebb\Bigl[K_r\Bigl(\norm{X_i - X_j}_2\Bigr)\Bigr] \\
& \leq M^2K_{\max}^3 p_{\max}^2 \cdot nr^{2 + 2d}
\end{align*}

\textit{Term 2.}
By symmetry, the same bound holds for term 2.

\textit{Term 3.}
Expressing $L_{n,r}f(X_i) = \sum_{k = 1}^{n} \bigl(f(X_i) - f(X_k)\bigr)K_r(X_i,X_k)$, by the linearity of expectation
\begin{align*}
\Ebb_{-i,j}\Bigl[L_{n,r}f(X_i)] & = (n - 2) L_{P,r}f(X_i) + \bigl(f(X_i) - f(X_j)\bigr)K_r(X_i,X_j) \\
\Ebb_{-i,j}\Bigl[L_{n,r}f(X_j)] & = (n - 2) L_{P,r}f(X_j) + \bigl(f(X_j) - f(X_i)\bigr)K_r(X_j,X_i)
\end{align*}
Upper bounding the square of sums by twice the sum of squares, we have
\begin{align*}
& \biggl(\bigl(\Ebb_{-i,j}\Bigl[\bigl(L_{n,r}f\bigr)(X_i) \Bigr] - \Ebb_{-i,j}\Bigl[\bigl(L_{n,r}f\bigr)(X_j)\Bigr]\biggr)^2 K_r\Bigl(\norm{X_i - X_j}_2\Bigr) \leq \\
&~~~ 2(n-2)^2\Bigl(\bigl(L_{P,r}f\bigr)(X_i) - \bigl(L_{P,r}f\bigr)(X_j)\Bigr)^2K_{r}(X_i,X_j) + 8 \bigl(f(X_j) - f(X_i)\bigr)^2 \bigl(K_r(X_j,X_i)\bigr)^3
\end{align*}
\section{Additional Theory}

We make some assumptions on the function $f$, the distribution $P$, and the kernel function $K$.

\begin{enumerate}[label=(A\arabic*)]
	\item 
	\label{asmp:smooth_function}
	$f \in C^{s}(\Xset;M)$ for some $s > 0$. If $s > 1$, then $f$ is also compactly supported on a strict subset of $\Xset$.

	\item 
	\label{asmp:smooth_density}
	$P$ admits a density $p$ with respect to the Lebesgue measure on $\Reals^d$. The density $p \in C^{k}(\Xset;p_{\max})$, for some $k > 0$.
	\item 
	\label{asmp:compact_kernel}
	$K$ is supported on a subset of $[0,1]$, and $K(z) \leq K_{\max} < \infty$ for all $z  \in [0,1]$.  
\end{enumerate}

Under these assumptions, we can bound various integrals.

\begin{lemma}
	\label{lem:expectation_1}
	Let $f$ satisfy~\ref{asmp:smooth_function} with $s = 1$, $P$ satisfy~\ref{asmp:smooth_density} with $k = 0$, and $K$ satisfy~\ref{asmp:compact_kernel}. Then for any $x \in \Xset$,
	\begin{equation*}
	\Ebb\Bigl[\bigl(f(x) - f(X)\bigr)^2 \bigl(K_r(\norm{x - X}_2)\bigr)^2\Bigr] \leq M^2 r^{2 + d} K_{\max}^2 p_{\max}
	\end{equation*}
\end{lemma}

\begin{lemma}
	\label{lem:expectation_2}
	Let $f$ satisfy~\ref{asmp:smooth_function} with $s = 2$, $P$ satisfy~\ref{asmp:smooth_density} with $k = 1$, and $K$ satisfy~\ref{asmp:compact_kernel}. Then for any $x \in \Xset$,
	\begin{equation*}
	\biggl|\Ebb\Bigl[\bigl(f(x) - f(X)\bigr) \bigl(K_r(\norm{x - X}_2)\bigr)\Bigr]\biggr| \leq M r^{2 + d} K_{\max} p_{\max}
	\end{equation*}
\end{lemma}

\begin{lemma}
	\label{lem:expectation_3}
	Let $f$ satisfy~\ref{asmp:smooth_function} with $s = 2$, $P$ satisfy~\ref{asmp:smooth_density} with $k = 1$, and $K$ satisfy~\ref{asmp:compact_kernel}. Then for any $x \in \Xset$,
	\begin{equation*}
	\biggl|\Ebb\Bigl[\bigl(f(x) - f(X)\bigr) \bigl(f(x) - f(X')\bigr) \bigl(K_r(\norm{x - X}_2)\bigr) \bigl(K_r(\norm{x - X'}_2)\bigr)\Bigr]\biggr| \leq M^2 r^{4 + 2d} K_{\max}^2 p_{\max}^2
	\end{equation*}
\end{lemma}

The following Lemma is more easily stated and proved using multi-index notation. For a function $f:\Reals^d \to \Reals$ which is $k$-times differentiable, and $\alpha \in [\mathbb{N}]^d$ satisfying $\abs{\alpha} := \alpha_1 + \ldots \alpha_d = k$,  we write
\begin{equation*}
D^{\alpha}f(x) = \frac{\partial^{k}f }{\partial x_1^{\alpha_1}\cdots\partial x_1^{\alpha_d}},~~ 
\end{equation*}
for the $(\alpha)$th-partial derivative of $f$ at $k$. Additionally, let $(x)^{\alpha} = x_1^{\alpha_1} \cdots x_d^{\alpha_d}$ for any $x \in \Reals^d$.

\begin{lemma}
	\label{lem:LB_approximation_1}
	Let $f$ satisfy~\ref{asmp:smooth_function} with $s > 2$, $P$ satisfy~\ref{asmp:smooth_density} with $k = 2$, and $K$ satisfy~\ref{asmp:compact_kernel}. Then for any $x \in \Xset$,
	\begin{equation*}
	content...
	\end{equation*}
\end{lemma}
\begin{proof}
	Let $\gamma = \min\{s - 2,1\}$. Since $f \in C^s(\Xset;M)$ for $s > 2$, we have that $f$ is twice-differentiable for all $x \in \Xset$, and moreover for any $y \in \Xset$,
	\begin{equation*}
	\biggl| f(y) - \Bigl(\sum_{\abs{\alpha} = 0}^2 D^{\alpha}f(x)(x)^{\alpha}\Bigr)  \biggr| \leq M \norm{y-x}_2^{\gamma}
	\end{equation*}
	Therefore,
	\begin{align*}
	-L_{P,r}f(x) & = \int \Bigl(f(y) - f(x)\Bigr)K_r(\norm{y - x}) \,dP(x) \\
	& = \sum_{\abs{\alpha} = 1}^{2} 
	\end{align*}
\end{proof}

\end{document}

### 2020-08-19 notes/weekly_notes/4.9.20.MeetingNotes.tex
Added weekly notes, which include some derivations of testing results for the fixed graph case.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\makeatletter
\newcommand{\vast}{\bBigg@{4}}
\newcommand{\Vast}{\bBigg@{5}}
\makeatother

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}
\newcommand{\Lap}{{\bf L}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\LE}{\mathrm{LE}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Meeting Notes for Week 4/3/20 - 4/9/20}
\author{Alden Green}
\date{\today}
\maketitle

Let $v_1,\ldots,v_n$ be eigenvectors of a graph $G$ (normalized so that $\norm{v_k}_n^2 = 1$), and $Y_1,\ldots,Y_n$ be observations on the vertices of $G$, with expectation $\Ebb[Y_i] = f_{0,i}$. The Laplacian eigenmaps estimator is $\wh{f}_{\LE} = \sum_{k = 1}^{\kappa} \dotp{Y}{v_k}_n v_k$. The bias of this estimator is 
\begin{equation*}
\Bigl\|\Ebb[\wh{f}_{\LE}] - f_0\Bigr\|_n^2 = \Bigl\|\sum_{k = 1}^{\kappa} \bigl(\dotp{f_0}{v_k}_n\bigr)^2 - f_0\Bigr\|_n^2 = \sum_{k = \kappa + 1}^{n} \bigl(\dotp{f_0}{v_k}_n\bigr)^2
\end{equation*}
Let $T_{\LE} = \norm{\wh{f}_{\LE}}_n^2$. The expectation of this statistic is
\begin{equation*}
\Ebb\bigl[T_{\LE}\bigr] = \frac{\kappa}{n} + \sum_{k = 1}^{\kappa} (\dotp{v_k}{f_0}_n)^2 = \frac{\kappa}{n} + \norm{f_0}_n^2 - \sum_{k = \kappa + 1}^{n} (\dotp{v_k}{f_0}_n)^2
\end{equation*}
Therefore in either the testing or estimation problems, we arrive at the same question: how to upper bound the tail sum $\sum_{k = \kappa + 1}^{n} (\dotp{v_k}{f_0}_n)^2$?

In our current analysis, we use the following upper bound, which is reminiscent of standard calculations involving the truncated series estimator:
\begin{align}
\sum_{k = \kappa + 1}^{n} (\dotp{v_k}{f_0}_n)^2 & \leq \frac{1}{\lambda_{\kappa + 1}^s} \sum_{k = \kappa + 1}^{n} \lambda_k^s (\dotp{v_k}{f_0}_n)^2 \nonumber \\
& \leq \frac{1}{\lambda_{\kappa + 1}^s} \sum_{k = 1}^{n} \lambda_k^s (\dotp{v_k}{f_0}_n)^2 \nonumber \\
& = \frac{f_0^T L^s f_0}{\lambda_{\kappa + 1}^s n} \label{eqn:approximation_error_bound}
\end{align} 
When the inner product $\dotp{v_k}{f_0}$ is taken with respect to a basis $v_k$ fixed a priori, the above calculation is tight: we can verify this by setting $f_0 \propto v_{\kappa + 1}$. However when the basis $v_{k}$ is random, setting $f_0 \propto v_{\kappa + 1}$ is no longer permissible, and we do not know whether or not the above calculation is tight. We investigate the matter empirically.

\section{Simulations}
In all of our simulations, $d = 1$. Each of the figures below examines the approximation error as function of $n$, for different values of the neighborhood graph radius $r$, and the smoothness of the function $f_0$. The regression function $f_0 = \sum_{k = 1}^{100} \theta_k \phi_k$ is constructed out of a basis $\phi_k$ of $H^s([0,1]^d)$, with $\theta_k$ chosen so that $|f_0|_{H^s} = 1$, and $|f_0|_{H^{q}} \gg 1$ for all $q > s$.

We plot four different quantities in each figure. The red line plots the actual approximation error $\sum_{k = \kappa + 1}^{n} (\dotp{v_k}{f_0}_n)^2$. The blue line plots our bound on approximation error, $(f_0^T L^s f_0)(\lambda_{\kappa + 1}^s n)$  The orange line plots the best approximation error we could hope for, $\sum_{k = \kappa}^{100} \theta_k^2$. The green line plots ${\kappa}^{-2s}$, the order of magnitude of the best approximation error (so the green and orange lines should track closely.) We usually choose $\kappa$ based on balancing bias and variance in the estimation context, but when things get interesting below we will see that the story is the largely the same for different values of $\kappa$.

\subsection{s = 1}

When $s = 1$ and $d = 1$, we have shown that 
\begin{equation*}
\frac{f_0^T L f_0}{\lambda_{\kappa + 1} n} \lesssim \kappa^{-2}
\end{equation*}
whenever $(\log(n)/n) \lesssim r \lesssim \kappa^{-1}$. Pictorially, this means that the blue line should be decreasing at least as fast as the orange and green lines. In Figures 1 we plot these lines as function of sample size $n$ on the log-log scale, making two different choices of $r \equiv (\log(n)/n)$ and $r \equiv n^{-1/3}$ (the left and right hand plots, respectively). Some algebra can verify that the latter choice satisfies $r \lesssim \kappa^{-1}$. We see that in both instances, all four lines track fairly closely together.

\begin{figure}[!ht]
	\centering
	\includegraphics[width = .48 \textwidth]{../../code/data/plots/approximation_plots/s1r1.pdf}
	\includegraphics[width = .48 \textwidth]{../../code/data/plots/approximation_plots/s1r2.pdf}
\end{figure}

\subsection{s = 2}

When $s = 2$ and $d = 1$, we have show that
\begin{equation*}
\frac{f_0^T L^2 f_0}{\lambda_{\kappa + 1}^2 n} \lesssim \kappa^{-4}
\end{equation*}
when $n^{-1/3} \lesssim r \lesssim \kappa^{-1}$. The choice $r \equiv n^{-1/3}$ still satisfies both of these conditions for our choice of $\kappa$, however clearly $r \equiv (\log(n)/n)$ violates the lower bound. Pictorially, this means we no longer have theoretical reason to believe that the blue line--and hence, the red line-- should decrease at least as fast as the orange and green lines. In Figures 2, we plot each of the four aforementioned lines as a function of sample size on the log-log scale. (The difference between the two rows is that in the second row, we take a larger value of $\kappa$; precisely, we multiply the previous value of $\kappa$ by a constant). In both Figures the conclusion is the same. When $r \equiv n^{-1/3}$ (right hand plot) all four lines still track reasonably close together. However, when $r \equiv (\log(n)/n)$ (left hand plot) the blue line no longer tracks the orange and green lines---in fact, it doesn't even decrease---but the red line still does.

\begin{figure}[!ht]
	\centering
	\includegraphics[width = .48 \textwidth]{../../code/data/plots/approximation_plots/s2r1.pdf}
	\includegraphics[width = .48 \textwidth]{../../code/data/plots/approximation_plots/s2r2.pdf}
	\includegraphics[width = .48 \textwidth]{../../code/data/plots/approximation_plots/s22r1.pdf}
	\includegraphics[width = .48 \textwidth]{../../code/data/plots/approximation_plots/s22r2.pdf}
\end{figure}

My conclusions are twofold: first, that our bound on $f_0^T L^s f_0/\lambda_{\kappa + 1}^s n$ appears tight, at least in these experiments, and second that the upper bound $f_0^T L^s f_0/\lambda_{\kappa + 1}^s n \geq \sum_{k = \kappa + 1}^{n} (\dotp{v_k}{f_0}_n)^2$ is not always tight.

\section{Simple example showing looseness of our bound.}

I will show that for a graph $\wb{G}_{n,r}$ which bears some resemblance to a neighborhood graph, $f_0$ have non-trivial loading onto many of the ``top'' eigenvectors, that is, eigenvectors which correspond to the largest eigenvalues in the graph Laplacian.

Let $X_1,\ldots,X_n$ be random samples which belong to $[0,1]^d$. Let $Z$ be the grid points spaced at interval width $r$; for $M = 1/r$,
\begin{equation*}
Z = \Bigl\{\frac{1}{M}(2m_1 - 1,\ldots,2m_d - 1): m \in [M]^d \Bigr\}.
\end{equation*}
and let $\wb{G}$ be the $d$-dimensional lattice graph over the grid points $X$.

Let $\Pi: \{X_1,\ldots,X_n\} \to Z$ assign each sample $X_i$ to its nearest neighbor $z \in Z$. Then $\wb{G}_{n,r} = ([n],E)$ will be the graph with edge set $E \subseteq [n] \times [n]$, with $(i,j)$ connected in $\wb{G}_{n,r}$ only if $(\Pi(X_i),\Pi(X_n))$ belongs to the edge set of $\wb{G}$.

It is an interesting fact---and not hard to show---that the eigenvectors of $\wb{G}_{n,r}$ can be divided into two categories. The first category consists of eigenvectors $v_k$ which are piecewise constant over cubes; that is, if $\Pi(X_i) = \Pi(X_j)$ then $v_{k,i} = v_{k,j}$. The second category consists of delta functions at points in the same cube: precisely, for each $i$ and $j$ such that $\Pi(X_i) = \Pi(X_j)$, the vector $u_k = e_i - e_j$ is an eigenvector of $\wb{G}_{n,r}$. Even if $f_0$ is smooth, for instance $f_0 \in C^s([0,1]^d)$ or $f \equiv \phi_{\kappa + 1}$, the loadings of $f$ onto the latter category of eigenvectors satisfy
\begin{equation*}
\dotp{f_0}{u_k}_n^2 = (f_{0,i} - f_{0,j})^2 \approx (\nabla f_0 (X_j))^2 \cdot r^2
\end{equation*}
which, while small, is not zero \emph{unless} $f_0$ itself happens to piecewise constant over cubes.

I don't really know how to use this information. If we were in fact interested in the graph $\wb{G}_{n,r}$, we might try to approximate $f_0$ by a piecewise constant function $\wb{f}_0$, then proceed to analyze the approximation error of $\wb{f}_0$ using~\eqref{eqn:approximation_error_bound}. But we aren't interested in $\wb{G}_{n,r}$. Nevertheless, I think the above example does give some useful insight into the empirical results we see in the previous section.

\end{document}

### 2020-08-19 notes/weekly_notes/6.18.19.NotesForWeek.tex
Added weekly notes, which include some derivations of testing results for the fixed graph case.

@@
 
+We show that a simple test based on the statistic $T_{\eta}$, with proper choice of the bandwidth parameter $r$, achieves minimax optimality over the class of Lipschitz functions $\mathcal{H}_{1}^{d}(L)$.
+
 \begin{theorem}
+	\label{thm: mmd_minimax_optimal}
+	Fix $r = n^{-2/(4+d)}$, and consider the test 
+	\begin{equation*}
+	\phi_{\eta,\tau} = \1\left\{ T_{\eta} \geq \tau \right\}, \quad \tau = a n^{-d/(4+d)}
+	\end{equation*}
+	with $a \geq 1$. There are constants $c_1$ depending only on $p_{\max}$, and $c_2$ depending only on $(L,p_{\max},p_{\min})$, such that
+	\begin{equation*}
+	R_{\epsilon}^{n}(\phi_{\eta,\tau}; \mathcal{H}_{1}^{d}(L)) \leq c_1/a^2, \quad \text{if}~ \epsilon \geq c_2 a n^{-2/(4 + d)}
+	\end{equation*}
+\end{theorem}
+\begin{proof}
+	Our first objective is to show that, for an appropriate choice of $c_1$, $\Ebb_{f,f}^{n}(\phi_{\eta,\tau}) \leq \frac{c_1}{2a^2}$ for all $f \in \mathcal{H}_{1}^{d}(L)$.
+This is equivalent to proving that
+	\begin{equation}
+	\label{eqn: mmd_minimax_optimal_2}
+	\Pbb \left(T_{\eta} \geq \tau \right) \leq c_1/2a^2
+	\end{equation}
+	which we now proceed to show.
 
-\end{theorem}
+	From Lemma \ref{lem: mmd_expectation}, we have that $\Ebb_{f,f}(T_{\eta}) = 0$. Therefore, by Chebyshev's inequality, 
+	\begin{equation}
+	\label{eqn: mmd_minimax_optimal_1}
+	\Pbb \left(T_{\eta} \geq \tau \right) \leq \Var(T_{\eta})/\tau^2
+	\end{equation}
+	In Lemma \ref{lem: var_mmd}, we have upper bounded $\Var(T_{\eta})$ as function of $r$ and $\delta^2(f,g)$.
+Under the null hypothesis $f = g$, $\delta^2(f,g) = 0$, and along with our choice of $r = n^{-2/(d + 4)}$, the upper bound on $\Var(T_{\eta})$ reduces to
+	\begin{equation*}
+	\Var(T_{\eta}) \leq \frac{p_{\max}^2}{2} n^{-2d/(d + 4)}
+	\end{equation*}
+	By our choice of $\tau = a n^{-d/(4 + d)}$, we therefore have
+	\begin{equation*}
+	\Var(T_{\eta})/\tau^2 \leq \frac{p_{\max}^2}{2a^2}
+	\end{equation*}
+	and therefore \eqref{eqn: mmd_minimax_optimal_2} holds with $c_1 := p_{\max}^2$.
+
+	We turn now to bound the type II error, $\Ebb_{f,g}^{(n)}(\phi_{\eta,\tau})$, for all $f,g \in \mathcal{H}_{1}^{d}(L)$ such that $\delta^2(f,g) \geq \epsilon = c_2 n^{-2/(4+d)}$.
+Our goal now will be to show
+	\begin{equation*}
+	\Pbb \left(T_{\eta} \leq \tau \right) \leq c_1/2a^2.
+	\end{equation*}
+	By Chebyshev's inequality we obtain
+	\begin{equation}
+	\label{eqn: mmd_minimax_optimal_4}
+	\Pbb(T_{\eta} \leq \tau) \leq \frac{\Var(T_{\eta})}{(\Ebb_{f,g}(T_{\eta}) - \tau)^2}
+	\end{equation}
+	We obtain the following lower bound on $\Var(T_{\eta})$ in Lemma \ref{lem: var_mmd} in terms of $r$ and $\delta(f,g)$:
+	\begin{equation*}
+	\Var(T_{\eta}) \leq 2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + \frac{p_{\max}^2}{2}r^d
+	\end{equation*}
+	We will split our analysis into cases, based on which term on the right hand side of the preceding equation dominates. \textcolor{red}{TODO: Complete this proof.}
+
+	\paragraph{Case 1, 1st term dominates:}
+	We assume
+	\begin{equation*}
+	2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) > \frac{p_{\max}^2}{2}r^d,
+	\end{equation*}
+	so that the variance is upper bounded
+	\begin{equation*}
+	\Var(T_{\eta}) \leq 4 n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right).
+	\end{equation*}
+	Additionally, in the following case, we will show that for a sufficiently large choice of $c_2$, depending on $p_{\max}, p_{\min}, L$ and $d$, we obtain
+	\begin{equation*}
+	\Ebb_{f,g} T_{\eta} \geq 2 \tau
+	\end{equation*}
+	so that 
+	\begin{equation*}
+	(\Ebb_{f,g} T_{\eta} - \tau)^2 \geq \left(\Ebb_{f,g} T_{\eta}\right)^2/4.
+	\end{equation*}
+	In Lemma \ref{lem: mmd_expectation}, moreover, we prove the following lower bound on $\Ebb_{f,g} T_{\eta}$ as a function of $r$ and $\delta(f,g)$:
+	\begin{equation*}
+	\Ebb_{f,g} T_{\eta} \geq \frac{(n - 1)p_{\min}^2}{ p_{\max}^2} \left(\delta^2(f,g)\frac{r^d}{2^d} - L\delta(f,g)r^{d + 1}\right).
+	\end{equation*}
+	We can therefore upper bound $\Var(T_{\eta})/(\Ebb_{f,g}T_{\eta} - \tau)^2$,
+	\begin{align*}
+	\Var(T_{\eta})/(\Ebb_{f,g}T_{\eta} - \tau)^2 \leq \frac{p_{\max}^6}{p_{\min}^6}  \frac{16 n r^{2d} \delta(f,g) \left(\delta(f,g) + Lr\right)}{(n-1)^2 r^{2d} \delta^2(f,g) \left(\delta(f,g){2^{-d}} - Lr\right)^2}
+	\end{align*}
+
+	\paragraph{Case 2, 2nd term dominates:}
+	We first lower bound the denominator, and then upper bound the numerator.
+In Lemma \ref{lem: mmd_expectation}, we prove the following lower bound on $\Ebb T_{\eta}$ when $f,g \in \mathcal{H}_1^{d}(L)$, stated as a function of $r$ and $\delta^2(f,g)$:
+	\begin{equation}
+	\label{eqn: mmd_minimax_optimal_3}
+	\Ebb_{f,g} T_{\eta} \geq \frac{(n - 1)p_{\min}^2}{ p_{\max}^2} \left(\delta^2(f,g)\frac{r^d}{2^d} - L\delta(f,g)r^{d + 1}\right).
+	\end{equation}
+	Given the choice of bandwidth parameter $r = n^{-2/(4+d)}$ and the lower bound $\delta(f,g) \geq c_2 a n^{-2/(d + 4)}$, we may further simplify the expression on the right hand side of \eqref{eqn: mmd_minimax_optimal_3},
+	\begin{align*}
+	\frac{(n - 1)p_{\min}^2}{ p_{\max}^2} \left(\delta^2(f,g)\frac{r^d}{2^d} - L\delta(f,g)r^{d + 1}\right) & = \frac{(n - 1)p_{\min}^2}{ p_{\max}^2} n^{-2d/(d+4)} \delta(f,g) \left(\frac{\delta(f,g)}{2^d} - Ln^{-2/(d + 4)} \right) \\
+	& \geq \frac{(n - 1)p_{\min}^2}{ p_{\max}^2} c_2 a n^{-(2d + 2)/(d+4)} \left(\frac{c_2 a n^{-2/(d + 4)}}{2^d} - Ln^{-2/(d + 4)} \right) \\
+	& \geq 2 a^2 n^{1 - (2d + 2)/(d + 4) - 2/(d + 4)} = 2 a^2 n^{-d/(d + 4)}.
+	\end{align*}
+	where the last inequality follows when $c_2$ is chosen to be sufficiently large relative to $p_{\min}, p_{\max}, L$ and $d$.
+Given our choice $\tau = a n^{-d/(d + 4)}$, the denominator of \eqref{eqn: mmd_minimax_optimal_4} can therefore be lower bounded
+	\begin{equation*}
+	(\Ebb_{f,g}(T_{\eta}) - \tau)^2 \geq (2a^2 - a)^2 n^{-2d/d(+4)} \geq a^4 n^{-2d/d(+4)} 
+	\end{equation*}
+	where the last inequality follows from the fact $2a^2 - a \geq a^2$ for all $a \geq 1$.
+
+	We turn now to upper bounding the numerator of \eqref{eqn: mmd_minimax_optimal_4}.
+In Lemma \ref{lem: var_mmd} we give the following upper bound on $\Var_{f,g} T_\eta$ as a function of $r$ and $\delta(f,g)$,
+	\begin{equation*}
+	\Var_{f,g} T_\eta \leq 2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + \frac{p_{\max}^2}{2}r^d 
+	\end{equation*}
+	Similar to before, we apply our choice of $r$ and lower bound on $\delta(f,g)$ and obtain
+	\begin{equation*}
+	2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + \frac{p_{\max}^2}{2}r^d \leq 
+	\end{equation*}
+\end{proof}
 
@@
 	\begin{equation*}
-	\Var(T_{\eta}) \leq 2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + 2r^d
+	\Var(T_{\eta}) \leq 2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + \frac{p_{\max}^2}{2}r^d
 	\end{equation*}
@@
 	\label{eqn: var_mmd_3}
-	\Ebb(\eta_r(X,Y)^2) = \Ebb(\eta_r(X,Y)) = r^d. 
+	\Ebb(\eta_r(X,Y)^2) = \Ebb(\eta_r(X,Y)) \leq \frac{p_{\max}^2}{4} r^d. 
 	\end{equation}
@@
 	\begin{equation*}
-	\Var(T_\eta) \leq 2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + 2r^d.
+	\Var(T_\eta) \leq 2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + \frac{p_{\max}^2}{2}r^d.
 	\end{equation*}

### 2020-08-19 notes/weekly_notes/6.9.19.NotesForWeek.tex
Added weekly notes, which include some derivations of testing results for the fixed graph case.

@@
 \newcommand{\dr}{\,dr}
+\newcommand{\wt}[1]{\widetilde{#1}}
 
@@
 \end{proof}
+
+\subsection{Relating the eigenvectors to a grid.}
+
+For $\kappa = n^{1/d}$, let $\widetilde{X} = \set{\wt{x} \in [\kappa]^d/\kappa}$ be a grid over the unit cube.
+Consider the lattice graph $\mathrm{Grid} = (\widetilde{V}, \widetilde{E})$, with vertices $\widetilde{V} = \set{v: v \in [\kappa]^d}$, and edges $\widetilde{E} = \set{(u,v):\widetilde{u}, \wt{v} \in \wt{V}, \norm{\wt{u} - \wt{v}}_1 = 1}$, and write $\widetilde{L} = \widetilde{V} \Lambda \wt{V}^T$ for the eigendecomposition of the associated Laplacian matrix.
+
+We wish to exhibit a mapping $P: \Reals^V \to \Reals^{\wt{V}}$ such that the difference between the statistics $T_{LE}^{(k)} = \ell^T V_k V_k^T \ell$ and $\wt{T}_{LE}^{(k)} = (P \ell)^T \wt{V}_k \wt{V}_k^T (P \ell)$ is small, which we accomplish through the use of optimal transport theory.
+
+\begin{lemma}
+	Fix $0 < \delta < 1$. For some number $c_1 > 0$, dependent on $f + g$, $d$, and $\delta$, but not $n$, with probability at least $1 - \delta$ there exists a mapping $T: X \to \wt{X}$ such that
+	\begin{equation}
+	\label{eqn: optimal_transport_to_grid}
+	\norm{X - TX}_{\infty} \leq c_1 \left(\frac{\log(n)}{n}\right)^{1/d}
+	\end{equation}
+	If $r = c_1 (\log n / n)^{1/d}$ and \ref{eqn: optimal_transport_to_grid} holds, then letting $P(v)$ be the index of $T(x_v)$ on the grid graph, we have for every $\theta \in \Reals^V$,
+	\begin{equation}
+	(P\theta)^T \wt{L} P \theta \leq \theta^T L \theta \leq (\log n)^{2 + 1/d} (P\theta)^T \wt{L} P \theta.
+	\end{equation}
+\end{lemma}
 

### 2020-08-19 notes/weekly_notes/8.7.19.NotesForWeek.tex
Added weekly notes, which include some derivations of testing results for the fixed graph case.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\domain}{\mathcal{X}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\Partial}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 8/7/19 - 8/15/19}
\author{Alden Green}
\date{\today}
\maketitle

Suppose we observe samples $X = (x_1, \ldots, x_n) \in \mathcal{X} \subseteq \Rd$. For a given $r > 0$, the $r$-neighborhood graph $G = (V,E)$ over $X$ is defined to be the undirected graph with vertex set $V = [n]$, and edge set $E = \set{(i,j): \norm{i - j}_2 \leq r}$. Let $A = (A_{ij})_{i,j = 1}^{n}$ be the adjacency matrix of $G$, with entries $A_{ij}$ equal to $1$ if $(i,j) \in E$ and $0$ otherwise. Let $D$ be the edge incidence matrix of $G$, with $\ell$th row $D_{\ell} = (0,\ldots,-1,\ldots,1,\ldots)$ with a $-1$ in the $i$th entry, and $1$ in the $j$th entry, and $0$ elsewhere, provided that the $\ell$th edge $e_{\ell} = (i,j)$ and $i < j$. Let $L = D^T D$ be the graph Laplacian matrix of $G$.

Let $\mathbb{Z}_{+}^d$ denote the set of all ordered $d$-tuples of nonnegative integers. For $\alpha \in \mathbb{Z}_{+}^d, \alpha = (\alpha_1,\ldots,\alpha_d)$, denote $\abs{\alpha} = \sum_{i = 1}^{d} \alpha_i$, and by $\mathcal{D}^{\alpha}f$ the partial derivative
\begin{equation*}
\Partial^{\alpha}f = \frac{\partial^{\abs{\alpha}}f}{\partial x_1^{\alpha_1}\cdots\partial x_d^{\alpha_d}}.
\end{equation*}
The Sobolev space $W^{s,p}(\mathcal{X})$ consists of all functions $f$ such that for each multiindex $\alpha$ with $\abs{\alpha} \leq k$, $D^{\alpha}f$ exists in the weak sense and belongs to $L^p(\mathcal{X})$. For $f \in W^{s,p}(\mathcal{X})$, we define the Sobolev norm of $f$ to be
\begin{equation}
\label{eqn:sobolev_norm}
\norm{f}_{W^{s,p}(\mathcal{X})} = \left(\sum_{\abs{\alpha} \leq s} \int_{\domain} \abs{D^{\alpha}f(x)}^p\right)^{1/p}.
\end{equation}
We will focus our attentions on $W^{s,2}(\domain)$.

For a function $f: X \to \Reals$, we can define a notion of smoothness of $f$ over $G$. Letting $\widetilde{f} = (f(x_1),\ldots,f(x_n)) \in \Rd$, introduce the smoothness functional
\begin{equation}
\label{eqn:discrete_sobolev_norm}
S(f) := \wt{f}^T L^{s} \wt{f}
\end{equation}
Our objective is to establish a relation between \eqref{eqn:sobolev_norm} and \eqref{eqn:discrete_sobolev_norm}. In particular, suppose the samples $x_i$ are sampled i.i.d from a probability measure $P$ over $\domain$. Our goal is to establish that, for an appropriate scaling $C_n$, with high probability
\begin{equation}
\label{eqn:continuous_to_discrete_embedding}
S(f) \leq C_n \norm{f}_{W^{s,2}(\mathcal{X})}
\end{equation}

\section{Expectation of $S(f)$.}

For $y \in \domain$, define the partial difference operator $D_y$ by $D_yf(x) = (f(y) - f(x))\eta(x,y)$, where $\eta(x,y) = \mathbf{1}(\norm{x - y} \leq r)$. For $\alpha = (\alpha_1,\ldots,\alpha_q) \in [n]^q$, define the iterated difference operator $D_{\alpha}$ to be
\begin{equation}
D_{\alpha}f(x) = D_{x_{\alpha_1}}(D_{x_{\alpha_2}}( \cdots D_{x_{\alpha_q}} f))(x)
\end{equation}

We begin by re-expressing the smoothness functional $S(f)$ in terms of sums of the iterated difference operator, similarly to \citep{sadhanala2017}.

\begin{lemma}
	\label{lem:smoothness_functional_difference_operator}
	For $s$ even, letting $q = s/2$, we have
	\begin{equation}
	\label{eqn:smoothness_functional_even}
	S(f) = \sum_{i = 1}^{n} \left(\sum_{\alpha \in [n]^q} D_{\alpha}f(x_i)\right)^2
	\end{equation}
	For $s$ odd, letting $q = (s-1)/2$, we have
	\begin{equation}
	\label{eqn:smoothness_functional_odd}
	S(f) = \frac{1}{2}\sum_{i, j = 1}^{n} \left(\sum_{\alpha \in [n]^q} D_{x_{i}}D_{\alpha}f(x_j))\right)^2
	\end{equation}
\end{lemma}
\begin{proof}
	Note that for any function $g: V \to \Reals$, $(-Lg)(x_i) = \sum_{j = 1}^{n} D_{x_j}g(x_i)$. To see this, note that $Dg$ is a length $m$ vector with $\ell$th entry $(Dg)_{\ell} = g_i - g_j$ provided that $e_{\ell} = (i,j)$ and $i < j$. Therefore, as $L g = \sum_{\ell = 1}^{m} D_{\ell} (D g)_{\ell}$, we have that $L g = \sum_{j > i} (g_i - g_j) \1((i,j) \in E) - \sum_{j < i} (g_j - g_i) \1((j,i) \in E) = \sum_{j = 1}^{n} (g_i - g_j) \eta_r(x_i,x_j) = \sum_{j = 1}^{n} D_{x_i}g(x_j)$. The statement follows since $-D_{x}f(y) = D_{y}f(x)$.

	Therefore, when $s$ is even, letting $q = s/2$, we have
	\begin{align*}
	f^T L^s f & = \sum_{i = 1}^{n} ((-L)^q f (x_i))^2 \\
	& = \sum_{i = 1}^{n} \left( \sum_{j_1 = 1}^{n} D_{x_{j_1}} L^{q-1}f (x_i) \right)^2.
	\end{align*} 
	If $q = 1$, this suffices. Otherwise, if $q \geq 2$, as $D_y$ is a linear operator, we obtain
	\begin{align*}
	\sum_{i = 1}^{n} \left( \sum_{j_1 = 1}^{n} D_{x_{j_1}} L^{q-1}f (x_i) \right)^2 & = \sum_{i = 1}^{n} \left( \sum_{j_1 = 1}^{n} D_{x_{j_1}}  \sum_{j_2 = 1}^{n} D_{x_{j_2}}L^{q-2}f (x_i) \right)^2 \\
	& = \sum_{i = 1}^{n} \left( \sum_{j_1j_2 = 1}^{n} D_{x_{j_1}} D_{x_{j_2}}L^{q-2}f(x_i) \right)^2
	\end{align*}
	and recursively, we arrive at the desired result.

	When $s$ is odd, letting $q = (s-1)/2$ we have that
	\begin{align*}
	f^T L^s f & = \sum_{i < j}^{n} \bigl((L^q f (x_{i}) - L^q f (x_{j})\bigr)^2 \1\bigl( (x_{i},x_{j}) \in E\bigr) \\
	& = \frac{1}{2} \sum_{i, j = 1}^{n} \bigl( (-L)^q f (x_{i}) - (-L)^q f (x_{j})\bigr)^2 \eta(x_{i}, x_{j}).
	\end{align*}
	By similar reasoning to above, we obtain
	\begin{equation*}
	f^T L^s f = \frac{1}{2} \sum_{i,j = 1}^{n} \left(\sum_{\alpha \in [n]^q} D_{\alpha}f(x_i)  - D_{\alpha}f(x_j)\right)^2\eta(x_i,x_j)
	\end{equation*} 
	Then, since $\eta = \eta^2$, we can replace $\eta(x_i,x_j)$ with $\eta^2(x_i,x_j)$ in the previous display, and obtain
	\begin{align*}
	f^T L^s f & = \frac{1}{2} \sum_{i,j = 1}^{n} \left(\sum_{\alpha \in [n]^q} D_{\alpha}f(x_i)  - D_{\alpha}f(x_j)\right)^2\eta^2(x_i,x_j) \\
	& = \frac{1}{2} \sum_{i,j = 1}^{n} \left(\sum_{\alpha \in [n]^q} \bigl(D_{\alpha}f(x_i)  - D_{\alpha}f(x_j)\bigr)\eta(x_i,x_j)\right)^2 \\
	\end{align*}
	and \eqref{eqn:smoothness_functional_odd} follows.
\end{proof}

\subsection{Expectation when $s = 2$.}

Let $P$ be absolutely continuous with density function $p$ over $\Rd$.

\begin{lemma}
	Suppose $f \in W^{2,2}(\mathcal{X})$, $p \in C^1(\domain)$, and $\partial X \in C^2$. Then there exists a constant $c$ which does not depend on $n$ or $f$ such that
	\begin{equation*}
	\Ebb(f^T L^2 f) \leq c \left(n^2 r^{d + 2} \sum_{\abs{\alpha} = 1} \norm{\Partial^{\alpha}f}_2^2  +  n^3 r^{2d + 4} \sum_{\abs{\alpha} \leq 2} \norm{\Partial^{\alpha}f}_2^2 \right)
	\end{equation*}
\end{lemma}
\begin{proof}
	Using Lemma \ref{lem:smoothness_functional_difference_operator}, we may rewrite
	\begin{align*}
	f^T L^2 f & = \sum_{i = 1}^{n} \left(\sum_{j = 1}^{n}D_{x_j}f(x_i)\right)^2 \\
	& = \sum_{i = 1}^{n} \sum_{j = 1}^{n} \sum_{k = 1}^{n} D_{x_j}f(x_i) D_{x_k}f(x_i) \\
	& = \sum_{i \neq j}^{n} \bigl(D_{x_j}f(x_i)\bigr)^2 + \sum_{i \neq j \neq k}^{n} D_{x_j}f(x_i) D_{x_k}f(x_i)
	\end{align*}
	so that the expectation $\Ebb(f^T L^2 f)$ becomes
	\begin{equation*}
	\Ebb(f^T L^2 f) = n(n-1) \Ebb\left(D_{X_1}f(X_2)\right)^2 + n(n-1)(n-2) \Ebb\left(D_{X_2}f(X_1) D_{X_3}f(X_2)\right).
	\end{equation*}

	The statement then follows from Lemmas \ref{lem:W22_penalty_operator_2} and \ref{lem:W22_penalty_operator_1}.
\end{proof}

\section{Additional Theory}

\begin{lemma}
	\label{lem:W22_penalty_operator_2}
	For any function f $\in W^{2,2}(\mathcal{X})$, if $\partial X \in C^2$ and $p \in C^1(\domain,L)$ then there exists a constant $c$ which does not depend on $f$ such that
	\begin{equation}
	\label{eqn:W22_penalty_operator_2}
	\Ebb(D(X_2)f(X_1)D_{X_3}f(X_1)) \leq c r^{2d + 4} \sum_{\abs{\alpha} \leq 2} \norm{\Partial^{\alpha}f}_{L^2}^2
	\end{equation}
\end{lemma}
\begin{proof}
	We rewrite the expectation on the right hand side of \eqref{eqn:W22_penalty_operator_2} as an integral:
	\begin{align*}
	\Ebb(D(X_2)f(X_1)D_{X_3}f(X_1)) & = \int_{\domain} \int_{\domain} \int_{\domain} D_yf(x) D_zf(x) \,dP(z) \,dP(y) \,dP(x) \\
	& = \int_{\domain} \left\{\int_{\domain} D_yf(x) \,dP(y)\right\}^2 \,dP(x) \\
	& \leq 2 \int_{\domain} \left\{\int_{\domain} D_yf(x) p(x) \,dy \right\}^2 +  \left\{\int_{\domain} D_yf(x) (p(y) - p(x)) \,dy \right\}^2 \,dP(x) \\
	& \leq 2 \int_{\Rd} \left\{\int_{\domain} D_yg(x) p(x) \,dy \right\}^2 +  \left\{\int_{\Rd} D_yg(x) (p(y) - p(x)) \,dy \right\}^2 \,dP(x)
	\end{align*}
	where $g$ is an extension of $f$ which is equal to $f$ almost everywhere on $\domain$, and satisfies $\norm{g}_{W^{2,2}(\Rd)} \leq c \norm{f}_{W^{2,2}(\Rd)}$ for some $c$ which does not depend on $f$. We will handle each term inside the integral separately. Assume without loss of generality that $g \in C^2$; otherwise, we can take $g_m \in C^2$ such that $\norm{g_m - g}_{W^{2,2}(\Rd)} \to 0$. For the first term, we use Lemma \ref{lem:local_average} to obtain
	\begin{equation*}
	\left\{\int_{\domain} D_yg(x) p(x) \,dy \right\}^2 \,dP(x) \leq \lambda_{\max}^3 \frac{d^2 r^{4 + 2d}}{\nu_d} \sum_{\abs{\alpha} = 2}\int_{B(0,1)} \int_{0}^{1} (1 - t)^2 \left(\Partial^{\alpha}g(x + trz)\right)^2 \,dt  \,dz,
	\end{equation*}
	and therefore,
	\begin{align*}
	\int_{\Rd} \left\{\int_{\domain} D_yg(x) p(x) \,dy \right\}^2 & \leq \lambda_{\max}^3 \frac{d^2 r^{4 + 2d}}{\nu_d} \sum_{\abs{\alpha} = 2}\int_{\Rd} \int_{B(0,1)} \int_{0}^{1} (1 - t)^2 \left(\Partial^{\alpha}g(x + trz)\right)^2 \,dt  \,dz \,dx \\
	& = \lambda_{\max}^3 \frac{d^2 r^{4 + 2d}}{\nu_d} \sum_{\abs{\alpha} = 2} \int_{B(0,1)}  \int_{0}^{1} (1 - t)^2 \int_{\Rd} \left(\Partial^{\alpha}g(x + trz)\right)^2 \,dx \,dt \,dz \\
	& = \frac{\lambda_{\max}^3 d^2 r^{4 + 2d}}{3} \sum_{\abs{\alpha} = 2} \norm{\Partial^{\alpha}g}_{L^2}^2.
	\end{align*}

	For the second summand, since $p(x) \in C^2(\domain; L)$, using reasoning similar to the proof of Lemma we have
	\begin{align*}
	\int_{\Rd} \left\{\int_{\Rd} D_yg(x) (p(y) - p(x)) \,dy \right\}^2 \,dP(x) & \leq \lambda_{\max} \int_{\Rd} \left(\int_{\Rd}( D_yg(x))^2 \,dy \right) \left(\int_{B(x,r)} (p(y) - p(x))^2 \,dy\right) \,dx \\
	& \leq \lambda_{\max} \int_{\Rd} \left(\int_{\Rd}( D_yg(x))^2 \,dy \right) \left(\int_{B(x,r)} L^2 \norm{x - y}^2 \,dy\right) \,dx \\
	& \leq \lambda_{\max} \nu_d r^{d + 2}\int_{\Rd} \int_{\Rd}( D_yg(x))^2 \,dy \,dx \\
	& \leq c \nu_d  \lambda_{\max} \nu_d r^{2d + 4} \sum_{\abs{\alpha} = 1} \norm{\Partial^{\alpha}f}_2^2.
	\end{align*}
\end{proof}

\begin{lemma}
	\label{lem:W22_penalty_operator_1}
	For any function $f \in W^{1,2}(\mathcal{X})$, if $\partial \domain \in C^1$, then there exists a constant $c$ which does not depend on $f$ such that
	\begin{equation*}
	\Ebb\left(D_{X_1}f(X_2)\right)^2 \leq c r^{d + 2} \sum_{\abs{\alpha} = 1} \norm{\Partial^{\alpha}f}_2^2
	\end{equation*}
\end{lemma}
\begin{proof}
	We write $\Ebb\left(D_{X_1}f(X_2)\right)^2$ as an integral,
	\begin{equation*}
	\Ebb\left(D_{X_1}f(X_2)\right)^2 = \int_{\domain} \int_{\domain} (f(x) - f(y))^2 \eta(y,x) \,dP(y) \,dP(x) \leq \lambda_{\max}^2 \int_{\domain} \int_{\domain} (f(x) - f(y))^2 \eta(y,x) \,dy \,dx
	\end{equation*}
	Now, as $\partial \domain \in C^{1}$, there exists a function $g:\Rd \to \Reals$ such that $g(x) = f(x)$, almost everywhere in $\domain$, and $\norm{g}_{W^{1,2}(\Rd)} \leq c \norm{f}_{W^{1,2}(\domain)}$.  Since $g = f$ almost everywhere in $\domain$, we have
	\begin{equation}
	\label{eqn:W12_integral_penalty_1}
	\int_{\domain} \int_{\domain} (f(x) - f(y))^2 \,dy \,dx \leq \int_{\Rd} \int_{\Rd} (g(x) - g(y))^2 \eta(y,x) \,dy \,dx.
	\end{equation}
	We may assume without loss of generality that $g \in C^{\infty}$ (otherwise take approximations $g_m \in C^{\infty}(\Rd)$, $g_m  \to g \in W^{s,2}(\Rd)$.) We expand
	\begin{align*}
	\int_{\Rd} (g(x) - g(y))^2 \,dy & = \int_{\Rd} \left( \int_{0}^{1} \frac{d}{dt} g(x + t(y - x)) \,dt \right)^2 \,dy \\
	& = \int_{\Rd} \left( \int_{0}^{1} \nabla g(x + t(y - x)) \cdot (y - x) \,dt \right)^2 \,dy \\
	& \leq \int_{\Rd} \int_{0}^{1} \norm{\nabla g(x + t(y - x))}^2 \norm{y - x}^2 \,dt \,dy
	\end{align*}
	Therefore,
	\begin{align*}
	\int_{\Rd} (g(x) - g(y))^2 \eta(y,x) \,dy \,dx & \leq \int_{\Rd} \int_{\Rd} \int_{0}^{1} \norm{\nabla g(x + t(y - x))}^2 \norm{y - x}^2 \,dt \eta(y,x) \,dy \,dx \\
	& \leq r^2 \int_{\Rd} \int_{\Rd} \int_{0}^{1} \norm{\nabla g(x + t(y - x))}^2 \,dt \eta(y,x) \,dy \,dx \\
	& = r^{2 + d} \int_{\Rd} \int_{B(0,1)} \int_{0}^{1} \norm{\nabla g(x + trz)}^2 \,dt \,dz \,dx  \tag{$z = (y - x)/r$}\\
	& =  r^{2 + d} \int_{B(0,1)} \int_{0}^{1} \int_{\Rd}  \norm{\nabla g(x + trz)}^2 \,dx \,dt \,dz \\
	& = \nu_d r^{2 + d} \sum_{\abs{\alpha} = 1} \norm{\Partial^{\alpha} g}_{L^2}^2 
	\end{align*}
\end{proof}

\begin{lemma}
	\label{lem:local_average}
	For any function $g \in C^2(\Rd)$, and any $x \in \Rd$
	\begin{equation*}
	\left(\int_{B(x,r)} g(y) - g(x) \dy \right)^2 \leq \frac{d^2 r^{4 + 2d}}{\nu_d} \sum_{\abs{\alpha} = 2}\int_{B(0,1)} \int_{0}^{1} (1 - t)^2 \left(\Partial^{\alpha}f(x + trz)\right)^2 \,dt  \,dz.
	\end{equation*}
\end{lemma}
\begin{proof}
	As $g \in C^2(\Rd)$, taking a Taylor expansion of $g$ around $x$, we obtain
	\begin{equation*}
	g(y) = g(x) + \sum_{\abs{\alpha} = 1 } \Partial^{\alpha}g(x) (y - x)^{\alpha} + \sum_{\abs{\alpha} = 2} (y - x)^{\alpha} \int_{0}^{1} (1 - t) \Partial^{\alpha} g(x + t(y - x)) \,dt.
	\end{equation*}
	We note that, by standard facts of the uniform distribution
	\begin{equation*}
	\int_{B(x,r)}(x-y)^{\alpha} = 0, \quad \textrm{for all $\abs{\alpha} = 1$}
	\end{equation*}	
	and we therefore have
	\begin{align*}
	\left(\int_{B(x,r)} g(y) - g(x) \dy \right)^2 & = \left(\int_{B(x,r)} \sum_{\abs{\alpha} = 2} (y - x)^{\alpha} \int_{0}^{1} (1 - t) \Partial^{\alpha} g(x + t(y - x)) \,dt\right)^2 \\
	& \leq \left(\int_{B(x,r)} \left(\sum_{\abs{\alpha} = 2} (y - x)^{2\alpha}\right)^{1/2} \left(\sum_{\abs{\alpha} = 2} \left\{\int_{0}^{1} (1 - t) \Partial^{\alpha} g(x + t(y - x)) \,dt\right\}^2 \right)^{1/2} \,dy \right)^2 \\
	& \leq d^2 r^{4} \left(\int_{B(x,r)} \left(\sum_{\abs{\alpha} = 2} \int_{0}^{1} (1 - t)^2 \left[\Partial^{\alpha} g(x + t(y - x))\right]^2 \,dt \right)^{1/2} \,dy \right)^2 \\
	& = d^2 r^{4 + 2d} \left(\int_{B(0,1)} \left(\sum_{\abs{\alpha} = 2} \int_{0}^{1} (1 - t)^2 \left[\Partial^{\alpha} g(x + trz)\right]^2 \,dt \right)^{1/2} \,dz \right)^2 \\
	& \leq \frac{d^2 r^{4 + 2d}}{\nu_d} \int_{B(0,1)} \sum_{\abs{\alpha} = 2} \int_{0}^{1} (1 - t)^2 \left[\Partial^{\alpha} g(x + trz)\right]^2 \,dt \,dz
	\end{align*}
\end{proof}

\bibliographystyle{plainnat}
\bibliography{../../graph_testing_bibliography}

\end{document}

### 2020-08-19 notes/weekly_notes/summer_progress.tex
Added weekly notes, which include some derivations of testing results for the fixed graph case.

--- 2019-11-23 notes/weekly_notes/11.15.19.NotesForWeek.tex
@@
 \begin{document}
-\title{Notes for Week 11/16/19 - 11/23/19}
+\title{Summary of Progress}
 \author{Alden Green}
@@
 
-In this week's notes, I pick up the thread of the ``11.8.19.NotesForWeek'' document.
-There, we obtained that the critical radius in the two-sample density testing problem over a graph $G$ was a function of
+A brief summary of the progress we've made so far.
+\section{Progress to date.}
+We observe samples $x_1,\ldots,x_n \sim P$ and $y_1,\ldots,y_m \sim Q$, (for convenience, assume $P$ and $Q$ are supported on $[0,1]^d$ absolutely continuous with respective densities $p$ and $q$ which are lower bounded away from $0$, and pper bounded away from $\infty$, over their support.) We started with the following IPM-based test statistic; construct a neighborhood graph $G$ with incidence matrix $D$, and let
 \begin{equation*}
-\Pi_{\max} = \max_{i = 1,\ldots,n} \set{\sum_{k = 1}^{\kappa} v_{k,i}^2}
+T_C := \sup_{\theta: \norm{D\theta}_2 \leq C} \dotp{\theta}{z},
 \end{equation*}
-where $L = V S V^T$ and $V = (v_1,\ldots,v_n)$ is an orthonormal matrix, $S$ is a diagonal matrix with increasing entries $0 = s_1 \leq s_2 \leq \ldots \leq s_n$.
-In particular, we showed that if $\Pi_{\max}$ could be upper bounded by a constant less than $1$, then the critical radius in the two-sample density testing problem over a graph $G$ was the same as in the one-sample regression testing problem.
-Our goal was therefore to upper bound $\Pi_{\max} \leq \frac{1}{2}$ for an appropriate choice of $\kappa$.
-We showed that such an upper bound holds when $G = G_{dn}$ the $d$-dimensional grid, and $\kappa \leq \frac{n}{4}$.
+where $z = 1_X - 1_Y$ is a contrast vector and $C$ is a scaling parameter which controls the size of the function class we compute the IPM over.
+We made the following alterations to the above setup and test statistic:
+\begin{itemize}
+	\item We modified the setup of the problem in several ways to ease tractability of analysis.
+First, we assumed we observed values $z_i = f(x_i) + \varepsilon_i$, where $x_i \sim (p + q)/2$, and $f = (p - q)/(p + q)$, and $\varepsilon_i$ represented the error of a Bernoulli random variable minus its expectation.
+Then, we let $\varepsilon_i$ be normally distributed.
 
-\section{Bounding $\Pi_{\max}$ for neighborhood graphs in $1$-dimension.}
-Now, suppose we additionally have a graph $\widetilde{G}$ with Laplacian matrix $\widetilde{L}$ such that
-\begin{equation}
-\label{eqn:spectral_similarity}
-(1 - \delta) x^T L x \leq x^T \widetilde{L} x \leq (1 + \delta) x^T L x, \quad \textrm{for some $0 \leq \delta \leq 1$ and all $x \in \Reals^n$.}
-\end{equation}
-Write the spectral decomposition $\widetilde{L} = U \Lambda U^T$, and suppose that the eigenvectors $U$ satisfy
-\begin{equation}
-\label{eqn:incoherence_tilde}
-u_{l,i}^2 \leq \frac{2}{n} \quad \textrm{for all $l,i = 1,\ldots,n$.}
-\end{equation}
+	\item More fundamentally, we changed the norm dictating the function class over which we take supremum.
+Let $C$ be such that if $f = (p - q) \in W^{2,1}(L)$, then $\norm{Df}_2 \lesssim C$.
+This choice of $C$ results in the IPM being computed over too rich of a function class, so must restrict the function class further.
+To do so, we modified the IPM as follows: letting $D^T D = L = V \Lambda V^T$ be the eigendecomposition of the Laplacian matrix $L$, let $g_C(\lambda) = \infty \cdot \1(\lambda > C)$, and let $g_C(L) = V g_C(\Lambda) V^T$.
+Then, consider the following choice of test statistic:
+	\begin{equation*}
+	T_C := \sup_{\theta: \theta^T g_C(L) \theta \leq C} \dotp{\theta}{z}
+	\end{equation*}
+	or equivalently
+	\begin{equation*}
+	T_C := \sum_{k: \lambda_k \leq C^2} \dotp{v_k}{z} 
+	\end{equation*}
+\end{itemize}
 
-\begin{lemma}
-	\label{lem:incoherence}
-	Let $\widetilde{G} = G_{1n}$ be the 1-dimensional grid graph (i.e.
-the chain), and assume $G$ satisfies \eqref{eqn:spectral_similarity}.
-Then we have the following bound on incoherence of the eigenvectors $v_{k,i}$: for any $R \gtrsim n \delta^{1/2} (s_{\max}^{1/2} \vee 1)$,
+For an appropriately formed neighborhood graph, and an appropriate choice of $C$, does $T_C$ achieve non-trivial power uniformly over a nonparametric function class , that is assume that
 \begin{equation*}
-	v_{k,i}^2 \lesssim \frac{1}{n}\left(R + \frac{n^4 \delta^2 s_{\max}^2}{R^3}\right) \quad \textrm{for any $i = 1,\ldots,n$}
-	\end{equation*} 
-	for any $k \leq n/8$.
-\end{lemma}
-
-Before proving Lemma~\ref{lem:incoherence}, we demonstrate its consequence.
-To state things concisely, we use tilde-order notation $\widetilde{O}$ and $\widetilde{\Theta}$ to suppress $\log(n)$ terms.
-Assume that
-\begin{equation}
-\label{eqn:neighborhood_graph_spectral_similarity}
-\textrm{$G = G_{n,r}$, $\widetilde{G} = G_{1n}$ satisfy \eqref{eqn:spectral_similarity} with $\delta = \widetilde{O}(n^{-1})$}, \quad \textrm{the maximum degree $d_{\max}(G) = \widetilde{O}(1)$}
-\end{equation}
-Using the upper bound on maximum eigenvalue of a Laplacian $s_{\max} \leq 2 d_{\max}$), by Lemma~\ref{lem:incoherence} the eigenvectors $v$ of $G$ satisfy
-\begin{equation*}
-v_{k,i}^2 = \widetilde{O}\biggl(\frac{1}{n}\left(R + \frac{n}{R^3}\right)\biggr)
-\end{equation*}
-for any $k \leq n/8$ and $R = \widetilde{\Theta}(n^{1/2}s_{\max}^{1/2}) = \widetilde{\Theta}(n^{1/2})$. Taking $R = \widetilde{\Theta}(n^{1/2})$, we have that
-\begin{equation}
-\label{eqn:incoherence}
-\max_{i = 1,\ldots,n} v_{k,i}^2 = \widetilde{O}\left(n^{-1/2}\right)
-\end{equation}
-for all $k \leq n/8$.
-This is quite a bit weaker than the incoherence bound~\eqref{eqn:incoherence_tilde} that the eigenvectors of the chain satisfy.
-Nevertheless, it will suffice for our purposes.
-When $\kappa \asymp n^{2/5}$, equation~\eqref{eqn:incoherence} implies that
-\begin{equation*}
-\Pi_{\max} = \widetilde{O}(\kappa n^{-1/2}) = o(1). 
-\end{equation*}
-We have previously shown that when $r = \widetilde{O}(n^{-1})$, the graph $G_{n,r}$ satisfies~\eqref{eqn:neighborhood_graph_spectral_similarity}  with high probability; therefore, $\Pi_{\max}$ is less than constant order.
-Looking back at the ``11.8.19.NotesForWeek'' (specifically equation (8) which gives the critical radius) we see that this is sufficient to ensure that the critical radius in the two-sample density testing and one-sample regression testing problems will be the same when $d = 1$.
-\section{Technical Results.}
-
-\subsection{Proof of Lemma~\ref{lem:incoherence}}
-
-The proof of Lemma~\ref{lem:incoherence} will proceed according to the following steps.
-\begin{enumerate}
-	\item We show that the incoherence $\max_{i} v_{k,i}^2$ can be upper bounded
-	\begin{equation}
-	\label{eqn:incoherence_proof_1}
-	\max_{i} v_{k,i}^2 \leq \frac{1}{n} \left(\sum_{j = 1}^{n} \abs{\dotp{v_k}{u_j}}\right)^2
-	\end{equation}
-	replacing the norm $\norm{v_k}_{\infty}$ by the inner products $\dotp{v_k}{u_j}$.
-	\item Let
-	\begin{equation*}
-	I(k,r) = \set{j \geq 0: \abs{j - k} \leq r}.
-	\end{equation*}
-	Using Davis-Kahan, we establish that the following upper bounds
-	\begin{equation}
-	\label{eqn:incoherence_proof_2}
-	\sum_{j \not\in I(k,r)}^{n} (\dotp{v_k}{u_j})^2 \leq 400 \frac{n^4 \delta^2 s_{\max}^2}{R^4}
-	\end{equation}
-	hold for each $k \leq n/8$ and any $r \geq 8k\sqrt{\delta}$. 
-	\item We carefully upper bound the $L_1$ norm present in \eqref{eqn:incoherence_proof_1} given the various bounds on $L_2$ norm we've established in \eqref{eqn:incoherence_proof_2}.
-\end{enumerate}
-
-\paragraph{Step 1: Proof of~\eqref{eqn:incoherence_proof_1}.}
-
-We can re-express $v_k$ in the basis $(u_j)$ as
-\begin{equation*}
-v_k = \sum_{j = 1}^{n} \dotp{v_k}{u_j} u_j,
-\end{equation*}
-whereupon the bound~\eqref{eqn:incoherence_proof_1} follows from \eqref{eqn:incoherence_tilde}. 
-\paragraph{Step 2: Proof of~\eqref{eqn:incoherence_proof_2}.}
-
-Let $\widetilde{L} = L + H$. By Davis Kahan, we have that
-\begin{equation}
-\label{eqn:davis_kahan}
-\sum_{j \not\in I(k,r)} (\dotp{v_k}{u_j})^2 \leq \left(\frac{\norm{H}_{\textrm{op}}}{\min{\abs{\lambda_j - s_k}:j \in I(k,r)}}\right)^2
-\end{equation}
-To upper bound the numerator, we use~\eqref{eqn:spectral_similarity} to get
-\begin{equation*}
-\norm{H}_{op} \leq \delta s_{\max}.
+\norm{f}_{W^{2,1}} := \sum_{\abs{\alpha} \leq 1} \norm{\Partial f}_{L^2}^2 \leq L
 \end{equation*}
 
-To lower bound the denominator, we note
-\begin{align}
-\abs{\lambda_j - s_k} & \geq \abs{\lambda_j - \lambda_k} - \abs{\lambda_k - s_k} \nonumber \\
-& \geq \abs{\lambda_j - \lambda_k} - \delta \lambda_k. \label{eqn:incoherence_proof_8}
-\end{align}
-The eigenvalues of the chain graph are well known to be
-\begin{equation*}
-\lambda_j = 2\left(1 - \cos\left(\frac{j\pi}{n}\right)\right) = 4\sin^2\left(\frac{k\pi}{n}\right) ~~\textrm{for $k = 0,\ldots,n - 1$.}
-\end{equation*}
-By Taylor expansion, for $k \leq j \leq n/(2\pi)$ we have
-\begin{equation}
-\label{eqn:incoherence_proof_5}
-\lambda_j - \lambda_k = 2\left(\cos\left(\frac{k\pi}{n}\right) - \cos\left(\frac{j\pi}{n}\right)\right) \geq \frac{(k - j)^2\pi^2}{n^2}.
-\end{equation}
-and when $j \geq n/(2\pi)$ we have $\lambda_j - \lambda_k \geq 2(\cos\left(\frac{\pi}{8}\right) - \cos\left(\frac{1}{2}\right) > .09$. Additionally since $\sin(x) \leq x$ for all $x \geq 0$, we have
-\begin{equation}
-\label{eqn:incoherence_proof_6}
-\lambda_k \leq 4\frac{k^2\pi^2}{n^2}.
-\end{equation}
-Combining~\eqref{eqn:incoherence_proof_5},\eqref{eqn:incoherence_proof_6}, and the lower bound $\abs{k - j} \geq R \geq 8k\sqrt{\delta}$, we have that
-\begin{equation}
-\label{eqn:incoherence_proof_7}
-\abs{\lambda_j - \lambda_k} - \delta \lambda_k \geq .04\frac{(k - j)^2\pi^2}{n^2} - 4 \delta\frac{k^2\pi^2}{n^2} \geq .02\frac{(k - j)^2\pi^2}{n^2}.
-\end{equation}
-The result follows from \eqref{eqn:incoherence_proof_7}, \eqref{eqn:incoherence_proof_8} and \eqref{eqn:davis_kahan}.
-
-\paragraph{Step 3: $L_1$ norm to $L_2$ norm.}
-
-Let $a \in \Reals^n$, and suppose we know that the $L_2$ norm of $a$ is bounded,
-\begin{equation}
-\label{eqn:incoherence_proof_3}
-\norm{a}_2^2 \leq 1.
-\end{equation}
-Under no other conditions on $a$, the upper bound $\norm{a}_1 \leq \sqrt{n}$ is the best that can be hoped for (achieved when $a = (n^{-1/2},\ldots,n^{-1/2})$).
-However, suppose we also know that for some $B_2^2 \geq B_3^2 \geq \ldots \geq B_n^2$, we have that there exists some $R \in [n]$ such that
-\begin{equation}
-\label{eqn:incoherence_proof_4}
-\sum_{j = r}^{n} a_j^2 \leq B_r^2 ~~\textrm{for each $r = R,\ldots,n$.}
-\end{equation}
-Clearly, if $B_n^2 \leq \frac{1}{n}$ then $a$ cannot be equal to $(n^{-1/2},\ldots,n^{-1/2})$.
-We might hope for more general improvements on the bound $\norm{a}_1 \leq \sqrt{n}$ if $B_R^2$ is quite small once $R$ gets sufficiently large.
-The following Lemma gives such a result.
-\begin{lemma}
-	\label{lem:incoherence_util_1}
-	Suppose $a \in \Reals^n$ satisfies \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4} for some $1 = B_1^2 = B_2^2 \geq \ldots \geq B_n^2 \geq B_{n+1}^2 = 0$.
-Assume additionally that there exists an $R \in [n]$ such that
-	\begin{equation}
-	\label{eqn:incoherence_util_1}
-	B_r^2 - B_{r + 1}^2 \leq \frac{1}{r}~\textrm{for each $r = R, R+1,\ldots, n - 1$.}
-	\end{equation}
-	Then,
-	\begin{equation}
-	\label{eqn:incoherence_util_2}
-	\norm{a}_1 \leq \sqrt{R}\sqrt{1 - B_R^2} + \sum_{j = R}^{n} \sqrt{B_j^2 - B_{j+1}^2}.
-	\end{equation}
-\end{lemma}
-\begin{proof}
-	Set
-	\begin{equation*}
-	(a_\star)_j^2 = 
-	\begin{cases*}
-	\frac{1}{R}(1 - B_R^2), ~ j \leq R \\
-	B_j^2 - B_{j + 1}^2, j > R
-	\end{cases*}
-	\end{equation*}
-	so that $\norm{a_\star}_1$ is equal to the right hand side of~\eqref{eqn:incoherence_util_2}.
-We now prove by contradiction that $a_\star$ maximizes $\norm{a}_1$ under the constraints \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4}.
-Suppose this were not true, and let $a$ be a vector such that $a$ satisfies \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4} and $\norm{a}_1 > \norm{a_{\star}}$.
-Then,
-	\begin{itemize}
-		\item since $\norm{a}_1 > \norm{a_{\star}}$ then there exists some index $j$ such that
-		\begin{equation}
-		\label{eqn:incoherence_util_3}
-		a_j > (a_{\star})_j
-		\end{equation} 
-		\item by~\eqref{eqn:incoherence_proof_3} $\norm{a}_2 \leq 1$. Since $\norm{a_{\star}}_2 = 1$, by \eqref{eqn:incoherence_util_3} there must exist some index $k$ such that
-		\begin{equation}
-		\label{eqn:incoherence_util_4}
-		(a_{\star})_k > a_k.
-		\end{equation} 
-		Choose $k$ to be the largest of all such indices.
-		\item suppose $k < j$ and $j > R$.
-Since $k$ was chosen to be the largest such index which satisfied~\eqref{eqn:incoherence_util_4}, clearly $a_i \geq (a_{\star})_i$ for all $i > j$, and by \eqref{eqn:incoherence_util_3} $a_j > (a_{\star})_j$.
-But then
-		\begin{equation*}
-		B_j^2 = \sum_{i = j}^{N} (a_{\star})_i^2 < \sum_{i = j}^{N} a_i^2 
-		\end{equation*}
-		and so $a$ does not satisfy~\eqref{eqn:incoherence_proof_4}. Therefore either $k > j$ or $j \leq R$.
-		\item In either case, we have
-		\begin{equation*}
-		a_j > (a_{\star})_j \geq (a_{\star})_k > a_k.
-		\end{equation*}
-		Let $j \leq l < k$ be the largest index for which $a_j > (a_{\star})_j$.
-
-		We now construct a vector $\norm{b}$ which satisfies the constraints \eqref{eqn:incoherence_proof_3} and \eqref{eqn:incoherence_proof_4} such that $\norm{b}_1 > \norm{a}_1$.
-Once we have shown this, we will have established a contradiction, and the proof of Lemma~\ref{lem:incoherence_util_1} will be complete.
-Let $b = (b_i)$ be given as follows:
-		\begin{equation*}
-		b_i = 
-		\begin{cases*}
-		a_i, ~~\textrm{if $i \neq k,l$}, \\
-		a_{\star,k}, ~~\textrm{if $i = k$}, \\
-		\sqrt{a_l^2 - (a_{\star,k}^2 - a_k^2)}, ~~\textrm{if $i = l$.}
-		\end{cases*}
-		\end{equation*}
-		Clearly $\norm{b}_2 = \norm{a}_2 = 1$ and so $b$ satisfies~\eqref{eqn:incoherence_proof_3}.
-To see that $b$ satisfies~\eqref{eqn:incoherence_proof_4} we separate the analysis into cases.
-When $r \geq l + 1$, we have that
-		\begin{equation*}
-		\sum_{i = r}^{n} b_i^2 = \sum_{i = r}^{n} a_i^2 \leq B_r^2.
-		\end{equation*}
-		When $l < r \leq k$, we have that
-		\begin{equation*}
-		\sum_{i = r}^{n} b_i^2 \leq \sum_{i = r}^{k} a_{\star,i}^2 + \sum_{i = k +1}^{n} a_{i}^2 \leq B_r^2 - B_{k + 1}^2 + B_{k + 1}^2 = B_r^2
-		\end{equation*}
-		Finally, when $R \leq r \leq l$, we have
-		\begin{equation*}
-		\sum_{i = r}^{n} b_i^2 \leq a_l^2 - (a_{\star,k}^2 - a_k^2) + (a_{\star,k}^2) + \sum_{i \neq k,l} a_i^2 = \sum_{i = r}^{n} a_i^2 \leq B_r^2.
-		\end{equation*}
-		Therefore $b$ satisfies~\eqref{eqn:incoherence_proof_4}. Finally, we have that
-		\begin{equation*}
-		\norm{b}_1 = \norm{a}_1 + (a_{\star,k}  - a_k) - (\sqrt{a_l^2 - (a_{\star,k}^2 - a_k^2)} - a_l) > \norm{a}_1 + (a_{\star,k}  - a_k) + (a_{\star,k}  - a_k) - (a_{\star,k}  - a_k) = \norm{a}_1,
-		\end{equation*}
-		so we have established the desired contradiction.
-	\end{itemize}
-\end{proof}
-
-\paragraph{Putting the pieces together.}
-We apply Lemma~\ref{lem:incoherence_util_1} to the vector $a = (a_i)_{i = 1}^{n}$, where
-\begin{equation*}
-a_i = \frac{1}{\sqrt{2}}\left(\abs{\dotp{v_k}{u_{k - i}}} + \abs{\dotp{v_k}{u_{k + i}}}\right).
-\end{equation*}
-Note the following:
-\begin{itemize}
-	\item $\norm{a}_2 \leq 1$.
-	\item By \eqref{eqn:incoherence_proof_2}, the vector $a$ satisfies the constraint~\eqref{eqn:incoherence_proof_4} with
-	\begin{equation*}
-	B_r^2 = 400\frac{n^4\delta^2s_{\max}^2}{r^4} ~~\textrm{when $r \geq 8 k \sqrt{\delta}$.}
-	\end{equation*}
-	\item Taylor expanding $f(r + 1) = (r + 1)^{-4}$ around $r$, we have that
-	\begin{equation*}
-	B_r^2 - B_{r + 1}^2 \leq 3200 \frac{n^4 \delta^2 s_{\max}^2}{r^5} ~~\textrm{when $r \geq 2$.}
-	\end{equation*}
-	Therefore, the sequence $(B_r)$ satisfies \eqref{eqn:incoherence_util_1} for all $r$ large enough such that 
-	\begin{equation*}
-	r \geq 2^{5/4} 100^{1/4} n \sqrt{\delta s_{\max}}
-	\end{equation*}
-	\item Since $a$ and $(B_r)$ satisfy the constraints \eqref{eqn:incoherence_proof_3}, \eqref{eqn:incoherence_proof_4} and \eqref{eqn:incoherence_proof_5}, we may apply Lemma~\ref{lem:incoherence_util_1} to $a$, obtaining that
-	\begin{equation*}
-	\sum_{j = 1}^{n} \abs{\dotp{u_j}{v_k}} = \sqrt{2}\norm{a}_1 \leq \sqrt{2R} + 60\sqrt{2} n^2 \delta s_{\max} \sum_{r = R}^{n} r^{-5/2}
-	\end{equation*}
-	for any $R \geq 2 \vee k\sqrt{\delta} \vee 2^{5/4} 100^{1/4} n \sqrt{\delta s_{\max}}$. Bounding sum by integral, we have that if $R \geq 2$ then
-	\begin{equation*}
-	\sum_{r = R}^{n} r^{-5/2} \leq \int_{R - 1}^{n - 1} x^{-5/2} \,dx \leq \frac{2^{3/2}2}{3} R^{-3/2}.
-	\end{equation*}
-	Plugging this into the previous expression, we arrive at
-	\begin{equation*}
-	\sum_{j = 1}^{n} \abs{\dotp{u_j}{v_k}} \leq \sqrt{R} + 60 \sqrt{2} n^2 \delta s_{\max} \frac{2^{3/2}2}{3R^{3/2}}
-	\end{equation*}
-	for any $R \geq 2 \vee k\sqrt{\delta} \vee 2^{5/4} 100^{1/4} n \sqrt{\delta s_{\max}}$. The desired result then follows from~\eqref{eqn:incoherence_proof_1}.
-\end{itemize}
-
-% \paragraph{Davis-Kahan sin-theta theorem.}
-%For a given $k \in [n]$ and $R \leq n$, decompose
-%\begin{equation*}
-%L = s_k v_k v_k^T + \sum_{k' \neq k}^{n} s_{k'} v_{k'} v_{k'}^T, ~~\textrm{and}~~ \widetilde{L} = \sum_{\ell:\abs{k - \ell} \leq R} \lambda_\ell u_\ell v_\ell^T + \sum_{\ell:\abs{k - \ell} > R} \lambda_\ell u_\ell u_\ell^T,
-%\end{equation*}
-%and denote $U_{k,R}^{\perp} = \sum_{\ell:\abs{k - \ell} > R} u_\ell u_\ell^T$.
-One consequence of the Davis-Kahan Theorem is the following: suppose for all $\ell: \abs{k - \ell} \geq R$, the eigenvalue $\lambda_{\ell}$ is excluded from the interval $(s_k - \delta,s_k + \delta)$ for some $\delta > 0$.
-Then
-%\begin{equation*}
-%\sum_{\ell:\abs{k - \ell} > R} \dotp{v_k}{u_{\ell}}^2 \leq \frac{\norm{(U_{k,R}^{\perp})^T H v_k}_F^2}{\delta^2}
-%\end{equation*}
-
 \end{document}

### 2020-08-19 notes/graph_regression2.tex

@@
 \DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
+\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}
 
@@
 \newcommand{\LS}{\mathrm{LS}}
+\newcommand{\SM}{\mathrm{SM}}
 \newcommand{\OS}{\mathrm{OS}}
@@
 
-In the random design nonparametric regression setup, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Rd$, and
+In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ with support $\Xset \subset \Rd$, and
 \begin{equation}
@@
 The two graph-based regression methods we will consider are \emph{Laplacian eigenmaps} and \emph{Laplacian smoothing}.
-Each use $\Lap_{n,r}$ to construct an estimator $\wh{f}$ of $f_0$.
+Each method uses $\Lap_{n,r}$ in a different manner to construct an estimator of $f_0$.
 Letting $v_1(G_{n,r}),\ldots,v_n(G_{n,r})$ denote eigenvectors of $\Lap_{n,r}$, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ \citep{belkin2003} is given by
@@
 In fact, as we will see in Section~\ref{sec:problem_setup_and_background} some of these methods are actually quite analogous to the graph Laplacian estimation and testing methods we have just introduced.
-Nevertheless, methods such as~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} deserve a place in the nonparametric regression toolbox.
+Nevertheless, methods such as~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} have a deserved place in the nonparametric regression toolbox.
 Their advantages include:
@@
 		\item Simplicity.
-		\item Computational efficiency.
-		\item Domain and density adaptivity.
+		\item Computational efficiency and stability.
+		\item Domain and density adaptivity. Easily generalizable. 
 		\item Practical success.
@@
 \textcolor{red}{(TODO): Include citations.} In this paper, we adopt a different perspective, and analyze~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} through the minimax lens.
-We show that for a variety of function classes $\mc{F}$, distributions $P$, and choices of domain $\mc{X}$ for which minimax rates are known, both~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are minimax optimal estimators (with loss measured in a suitable norm) and both~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} are minimax optimal test statistics.
-This lends additional theoretical support to the case for using graph Laplacian methods.
-	\item \textbf{Summary}
+We show that when the regression function $f_0$ is suitably smooth (by which we mean that it belongs to a first-order Sobolev ball; roughly speaking, that it possesses weak derivative $Df$ bounded in $L^2(\Xset)$ norm) both~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are minimax optimal estimators, and both~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} are minimax optimal test statistics.
+We also show that under appropriate conditions, these estimators and test statistics can leverage additional assumed structure, either on the regression function $f_0$ or distribution $P$.
+	\item \textbf{Organization.}
+\end{itemize}
+
+\paragraph{Notation.}
+\begin{itemize}
+	\item We write $\mathbb{Z}^d$ for $\mathbb{Z} \otimes \cdots \mathbb{Z}$, where the tensor product is taken a total of $d$ times.
+	\item We denote by $\Leb^2(\Xset)$ be the set of square integrable functions $f:\Xset \to \Reals$.
+	\item For functions $f,g$ which map from $\Xset$ to $\Reals$, we write $\dotp{f}{g}_n := \frac{1}{n}\sum_{i = 1}^{n} f(X_i) g(X_i)$, which we refer to as the \emph{empirical inner product}, and $\norm{f}_n^2 = \dotp{f}{f}_n$, which we refer to as the (squared) \emph{empirical norm}.
 \end{itemize}
@@
 
-Before we get to our main results, we will review some of the relevant literature on nonparametric estimation and regression, and recall some relevant minimax analysis.
-We will also briefly recall some (of the many) estimators and tests which achieve these rates.
-In particular, we will focus on the \emph{truncated series} and \emph{smoothing spline} methods.
-As we will see, Laplacian eigenmaps and Laplacian smoothing can be seen as graph-based analogues to these ``continuuum'' methods.
+Before we get to our main results, we will review some of the relevant literature on minimax optimal nonparametric estimation and testing of Sobolev functions.
+We will also briefly recall some (of the many) estimators and tests which achieve these rates: in particular the \emph{truncated series} and \emph{smoothing spline} methods.
+As we will see, Laplacian eigenmaps and Laplacian smoothing can be seen as graph-based, i.e.
+discrete, analogues to these ``continuuum'' methods.
+
+\subsection{Sobolev spaces}
+\label{subsec:sobolev_spaces}
 
 \begin{itemize}
-	\item \textbf{Minimax estimation.} For given integers $s > 0$ and $p > 0$, the Sobolev space $W^{s,p}(P)$ consists of all functions $f \in \Leb^p(P)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d)$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^p(P)$.
-The Sobolev $\{s,p\}$ norm is then
+	\item \textbf{Sobolev spaces.}  Roughly speaking, Sobolev spaces contain functions $f \in \Leb^p(\Xset)$ with derivatives which themselves belong to $\Leb^p(\Xset)$.
+More formally, for given integers $s, p > 0$, the Sobolev space $W^{s,p}(\Xset)$ consists of all functions $f \in \Leb^p(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^p(\Xset)$.
+The Sobolev-$\{s,p\}$ seminorm $\seminorm{f}_{W^{s,p}(\Xset)}$ and norm $\norm{f}_{W^{s,p(\Xset)}}$ are then given by
 	\begin{equation*}
-	\norm{f}_{W^{s,p}(P)}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^p \,dP
+	\seminorm{f}_{W^{s,p}(\Xset)}^p = \sum_{\abs{\alpha} = s}\int_{\mathcal{X}} \Bigl|\bigl(D^{\alpha}f\bigr)(x)\Bigr|^p \,dx, ~~ \norm{f}_{W^{s,p}(\Xset)}^p = \sum_{k = 0}^{s} \seminorm{f}_{W^{k,p}(\Xset)}^p
 	\end{equation*}
-	and for a given $M > 0$, the corresponding ball is $W^{s,p}(P, M) = \set{f: \norm{f}_{W^{s,p}(P)} \leq M}$.
-In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space; we adopt the usual convention of writing $H^s(P) = W^{s,p}(P)$ and $H^s(P,M) = W^{s,2}(P,M)$.
+	and for a given $M > 0$, the corresponding ball is $W^{s,p}(\Xset, M) = \set{f: \norm{f}_{W^{s,p}(\Xset)} \leq M}$.
+In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space, and we adopt the usual convention of writing $H^s(\Xset) = W^{s,p}(\Xset)$ and $H^s(\Xset,M) = W^{s,2}(\Xset,M)$.
 We will confine our attention to this special case hereafter.
-Let us also make the following assumption:
-	\begin{enumerate}[(P1)]
+	\item \textbf{Connections to sequence spaces.} The balls $H^s(\Xset,M)$ are closely connected with the \emph{Sobolev ellipsoids} $\Theta_s(M) \subseteq \ell^2(\mathbb{Z}^d)$, defined as
+	\begin{equation*}
+	\Theta_s(M) = \Bigl\{\theta \in \ell_2(\mathbb{Z}^d): \sum_{k \in \mathbb{Z}^d} \theta_k^2 \abs{k}^{2s} < M\Bigr\}
+	\end{equation*}
+	Let us assume that the domain $\Xset = [0,1]^d$. Letting be $\set{\phi_k}_{k \in \mathbb{Z}^d}$ be the tensor product trigonometric polynomial basis of $\Leb^2(\Xset)$, the space
+	\begin{equation*}
+	H_{\textrm{per}}^s(\Xset,M) = \Bigl\{\sum_{k \in \mathbb{Z}^d} \theta_k \phi_k: \theta_k \in \Theta_s(M) \Bigr\}
+	\end{equation*}
+	satisfies $H_{\textrm{per}}^s(\Xset,M) \subseteq H^s(\Xset,M')$ for a sufficiently large value of $M_1 > 0$.
+On the other hand, any $1$-periodic function $f \in W^{s,p}(\Xset,M)$ also belongs to $H_{\textrm{per}}^s(\Xset,M_0)$ for a sufficiently large value of $M_0 > 0$.
+	\item \textbf{Zero-trace Sobolev spaces.} We will also be interested in the \emph{zero-trace} Sobolev spaces.
+\end{itemize}
+
+\subsection{Orthogonal series and smoothing splines.}
+\label{subsec:continuum_methods}
+When the regression function is assumed to belong to a Sobolev space, two approaches particularly well-suited for nonparametric regression are orthogonal series \textcolor{red}{projection}, and smoothing splines.
+The orthogonal series estimator $\wh{f}_{\OS}$ is given by
+\begin{equation}
+\label{eqn:orthogonal_series}
+\wh{f}_{\OS} = \sum_{k: \abs{k} \leq \kappa} \dotp{Y}{\phi_k}_n \phi_k .
+\end{equation}
+where $\kappa \geq 0$ is a hyperparameter, and we recall that $\{\phi_k\}_{k \in \mathbb{Z}^d}$ is the tensor product trigonometric polynomial basis of $\Leb^2(\Xset)$.
+The smoothing spline estimator $\wt{f}_{\SM}$ is a variational estimator, defined as the solution to the following optimization problem,
+\begin{equation}
+\label{eqn:smoothing_spline}
+\wt{f}_{\SM} = \argmin_{f} \sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \seminorm{f}_{H^{s}(\Xset)}^2
+\end{equation}
+where the minimum is taken over all $f \in H^s(\Xset)$.
+
+\begin{itemize}
+	\item \textbf{Connecting orthogonal series and Laplacian eigenmaps estimators.}The orthogonal series estimator $\wh{f}_{\OS}$ is closely connected with the Laplacian eigenmaps estimator $\wh{f}_{\LE}$.
+At a high-level, one can go from~\eqref{eqn:orthogonal_series} to~\eqref{eqn:laplacian_eigenmaps} simply by substituting $\{v_k\}_{k \in [n]}$ for $\{\phi_k\}_{k \in \mathbb{Z}^d}$.
+More deeply, just as the vectors $v_k$ are eigenvectors of the graph Laplacian $\Lap_{n,r}$, so too are the functions $\phi_k$ eigenfunctions of a particular differential operator $\Delta: C^2(\Xset) \to \Leb^2(\Xset)$, known as the \emph{Laplace} operator and formally defined as
+	\begin{equation}
+	\label{eqn:laplace_operator}
+	\Delta g := -\mathrm{div}(\nabla g).
+	\end{equation}
+	In fact, it has been shown in \textcolor{red}{(references)} that when the radius $r \to 0$ as $n \to \infty$, the graph Laplacian $\Lap_{n,r}$ converges in an appropriate sense to a weighted version $\Delta_{P}$ of $\Delta$, with
+	\begin{equation}
+	\Delta_{P} g := -\frac{1}{p}\mathrm{div}(p^2 \nabla g).
+	\end{equation}
+	Along similar lines, \textcolor{red}{(references)} have established that for any fixed $k \geq 0$, as $n \to \infty$ the eigenvector $v_k$ will converge, again in an appropriate sense, to an eigenfunction $\phi_k$ of $\Delta_{P}$.
+The upshot is that when $P$ is the uniform measure, for an appropriate sequence of radii $r(n) \to 0$ and any fixed $\kappa \geq 0$ the estimator $\wh{f}_{OS}$ is in fact the limit of $\wh{f}_{\LE}$, in the sense that
+	\begin{equation}
+	\label{eqn:laplacian_eigenmaps_consistency}
+	\abs{\wh{f}_{\LE}(X_i) - \wh{f}_{\OS}(X_i)} \overset{P}{\to} 0~~\textrm{as $n \to \infty$.}
+	\end{equation} 
+	for each $i \in \mathbb{N}$.
+	\item 
+	\textbf{Connecting smoothing spline and Laplacian smoothing estimators.} In a similar vein, the smoothing spline estimator $\wh{f}_{\SM}$ has deep connections with the Laplacian smoothing estimator $\wh{f}_{\LS}$.
+Using the definition~\eqref{eqn:laplace_operator}, recursively applying integration by parts, and then invoking the aforementioned convergence properties of $\Lap_{n,r}$, it can be shown that \textcolor{red}{references} that for any $f \in C^{2s}(\Xset)$,
+	\begin{equation}
+	\label{eqn:pointwise_seminorm_convergence}
+	\abs{f}_{H^s(\Xset)} = \int_{\Xset} \bigl(\Delta^sf\bigr)(x) f(x) \,dx = \lim_{n \to \infty} f^T \Lap_{n,r}^s f.
+	\end{equation}
+	By applying appropriation concentration inequalities, one can extend~\eqref{eqn:pointwise_seminorm_convergence} to hold uniformly over all $f \in H^{2s}(\Xset)$.
+While this does not imply a formal statement about the convergence of $\wh{f}_{\LS}$ to $\wh{f}_{SM}$ as in~\eqref{eqn:laplacian_eigenmaps_consistency}---note the difference between the domain of minimization in~\eqref{eqn:smoothing_spline} and the regularity required for~\eqref{eqn:pointwise_seminorm_convergence}---it nevertheless makes clear that the estimators $\wh{f}_{\LS}$ and $\wh{f}_{\SM}$ are linked.
+	\item \textbf{Domain and density adaptivity.} \textcolor{red}{(TODO)}
+\end{itemize}
+
+\subsection{Minimax rates over Sobolev classes}
+\label{subsec:minimax_rates}
+One reason all of this is intriguing is that, as we mentioned previously, the continuum estimators $\wh{f}_{\OS}$ and $\wh{f}_{\SM}$---and test statistics based on functionals of these estimators--- have strong statistical properties.
+In particular, these estimators and tests are known to have error rates within a constant factor of (minimax) optimal.
+
+\paragraph{Estimation rates.}
+Under appropriate regularity conditions, the minimax estimation rate over Sobolev balls is 
+\begin{equation}
+\label{eqn:sobolev_space_estimation_minimax_rate}
+\inf_{\wh{f}} \sup_{f_0 \in \mc{H}^s(M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d \geq 0$,}
+\end{equation}
+where the infimum is over all measurable maps $\wh{f}(X_1,Y_1,\ldots,X_n,Y_n)$, and the function class $\mc{H}^s(M)$ is one of $H^s(\Xset,M)$, $H_{\mathrm{per}}^s(\Xset,M)$ or $H_0^s(\Xset,M)$.
+Orthogonal series and smoothing spline estimators are two (of the many) estimators which certify the upper bound in~\eqref{eqn:sobolev_space_estimation_minimax_rate}.
+Concretely, if we are willing to assume
+\begin{enumerate}[label=(P\arabic*)]
+	\item
+	\label{asmp:uniform_density}
+	the distribution $P$ is the uniform probability measure defined on domain $\Xset = [0,1]^d$.
+\end{enumerate}
+then the orthogonal series estimator is minimax optimal over the Sobolev ellipsoids $H_{\mathrm{per}}^s(M)$, meaning
+\begin{equation*}
+\sup_{f \in H_{\mathrm{per}}^s(\Xset,M)}\Ebb\Bigl[\norm{\wh{f}_{\OS} - f}_{L^2(\Xset)}^2\Bigr]  \lesssim n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d \geq 0$.}
+\end{equation*}
+An advantage of $\wt{f}_{\SM}$ over $\wh{f}_{\OS}$ is that it requires less stringent assumptions on the design density in order to be minimax optimal.
+Instead of Assumption~\ref{asmp:uniform_density}, we will merely assume that the density is bounded above and below.
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{1}
 	\item 
@@
 	The distribution $P$ is defined on domain $\Xset = [0,1]^d$, and admits a density $p$ with respect to the uniform (Lebesgue) measure on $[0,1]^d$.
-The density $p$ is bounded away from $0$ and $\infty$,
+The density $p$ is bounded away from $0$ and $\infty$, that is
 	\begin{equation*}
-		0 < p_{\min} < p(x) < p_{\max} < \infty
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty
 	\end{equation*}
@@
 \end{enumerate}
-	The assumption~\ref{asmp:bounded_density} is a standard but significant assumption; we will return to discuss it further at \textcolor{red}{a later point}.
-Assuming~\ref{asmp:bounded_density}, the minimax estimation rate over the Sobolev ellipsoids $H^{s}(P,M)$ is
+The assumption~\ref{asmp:bounded_density} is itself a standard but significant assumption; we will return to discuss it further at \textcolor{red}{a later point}.
+Assuming~\ref{asmp:bounded_density}, the estimator $\wt{f}_{\SM}$ satisfies
 \begin{equation*}
-	\inf_{\wh{f}} \sup_{f \in H^s(P,M)} \Ebb\Bigl[\norm{\wh{f} - f}_{L^2(P)}^2\Bigr] \asymp n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d \geq 0$,}
+\sup_{f_0 \in H^s(\Xset,M)}\Ebb\Bigl[\norm{\wh{f}_{\SM} - f_0}_{L^2(\Xset)}^2\Bigr]  \lesssim n^{-2s/(2s + d)}~~\textrm{for all $2s > d$.}
 \end{equation*}
-	where the infimum is over all measurable maps $\wh{f}(X_1,Y_1,\ldots,X_n,Y_n)$
-	\item \textbf{Minimax hypothesis testing.}
-	\item \textbf{Our contributions, long version.}
+The restriction $2s > d$ is a severe limitation of smoothing splines relative to  orthogonal series.
+It is worth emphasizing that this is a fundamental defect of the smoothing spline estimator, as opposed to any flaw in the theory; when $2s \leq d$, the optimization problem~\eqref{eqn:smoothing_spline} is not even well-defined!
+\textcolor{red}{(TODO):} succinctly explain why.
+
+\paragraph{Goodness-of-fit testing rates.}
+
+\begin{itemize}
+	\item \textbf{Minimax hypothesis testing.} In the nonparametric regression hypothesis testing problem, we ask for a test function (formally, a Borel measurable map $\phi: \mc{X}^n \times \Reals^n \to \{0,1\}$) which can distinguish between the null and alternative hypotheses
+	\begin{equation}
+	\mathbf{H}_0: f_0 = 0, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \neq 0.
+	\end{equation}
+	We measure the performance of a test $\phi$ by its Type I error $\Ebb_0(1 - \phi)$, and Type II error $\mathbb{E}_{f_0}(\phi)$.
+It is well known that no test $\phi$ with controlled Type I error---say, $\Ebb_0(\phi) \leq 1/b$ for some constant $b > 1$---
+	can have controlled Type II error---$\Ebb_{f_0}(\phi) \leq 1/b$--- for every $f_0 \neq 0$.
+As a result, it is standard to impose two kinds of restrictions on the alternative hypothesis class $\mathbf{H}_a$.
+The first is a smoothness condition on $f_0$: specifically, we will assume that $f_0$ belongs to a Sobolev ball $\mc{H}^s(M)$.
+The latter is a separation condition, enforcing a minimum distance between $f_0$ and $0$: specifically, we will insist that $\norm{f_0  - 0}_{\Leb^2(\Xset)} > \epsilon$ for some $\epsilon > 0$.
+	\item\textbf{Minimax testing rates over Sobolev classes.} Once smoothness and separation conditions are established, the following becomes a natural question to ask: what is the smallest value of $\epsilon$ for which there exists a test $\phi$ with both small Type I error, and small Type II error over all functions $f_0$ which are smooth and separated from $0$ by at least $\epsilon$?
+This smallest separation $\epsilon$ is known as the critical radius, and in the case of the Sobolev classes it is given by
+	\begin{equation*}
+	\epsilon_n^{\star}\bigl(\mc{H}^s(M)\bigr):= \inf\biggl\{\epsilon: \inf_{\phi} \Bigl\{\Ebb_0(\phi) +  \sup_{\substack{f_0 \in \mc{H}^s(M) \\ \norm{f_0}_{\Leb^2(\Xset)} \geq \epsilon}} \Ebb_{f_0}[1 - \phi]\Bigr\} \leq \frac{2}{b} \biggr\} \asymp n^{-2s/(4s + d)}
+	\end{equation*}
+	for any $4s > d$, and $\mc{H}^s(M)$ one of $H^s(\Xset,M)$, $H_{\mathrm{per}}^s(\Xset,M)$ or $H_0^s(\Xset,M)$. On the other hand, when $4s \leq d$ the minimax rate is dimension-free,
+	\begin{equation*}
+	\epsilon_n^{\star}\Bigl(\mc{H}^s(M)\Bigr) \asymp n^{-1/4},
+	\end{equation*}
+	revealing a distinction between the estimation and goodness-of-fit testing problems.
+	\item \textbf{Error rates of orthogonal series and smoothing spline-based tests.} 
+	Just as $\wh{f}_{\OS}$ and $\wh{f}_{\SM}$ achieve minimax optimal estimation rates, tests using the plug-in statistics
+	\begin{equation*}
+	T_{\OS} = \bigl\|\wh{f}_{\OS}\bigr\|_{\Leb^2(\Xset)}^2,~~\textrm{and}~~T_{\SM} = \bigl\|\wt{f}_{\SM}\bigr\|_{n}^2
+	\end{equation*}
+	are minimax optimal for the goodness-of-fit testing problem over Sobolev balls (see \citet{ingster2009} for analysis of the former statistic, and \citet{liu2019} for analysis of the latter).
+\end{itemize}
+
+\subsection{Our contributions}
+\label{subsec:our_contributions}
+
+\begin{itemize}
+	\item \textbf{Departure from what's already known.} At a high level, the connections between the graph-based methods~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} and the continuum methods outlined in Subsection~\ref{subsec:continuum_methods} give reason to believe that the strong statistical properties of the latter are shared by the former.
+Making this rough statement precise is the content of this paper.
+In particular, we will prove the following results:
 	\begin{itemize}
 		\item \textbf{Minimax optimal estimation.}
-		We show that with high probability,  $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2s/(2s + d)}$ whenever $f_0 \in C^s(\mc{X})$ for any $s \leq 1$, or $f_0 \in H^s(P)$ for $s = 1$.
+		We show that with high probability, $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2s/(2s + d)}$ over all $f_0 \in H^1(\Xset;M)$.
 The same conclusions hold with respect to $\wt{f}_{\LS}$ when $d \leq 4$.
 		\item \textbf{Minimax optimal testing.}
-		We show that with high probability, a level-$\alpha$ test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in C^s(\mc{X})$ for any $s \leq 1$---or $f_0 \in H^s(P)$ for $s = 1$---additionally satisfies $\norm{f_0}_{\Leb^2(P)} \gtrsim \max\{n^{-4s/(4s + d)},n^{-1/4}\}$.
-\textcolor{red}{The same conclusions hold with respect to $\wh{f}_{LS}$.}
+		We show that with high probability, a test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in H^1(\Xset;M)$ satisfies $\norm{f_0}_{\Leb^2(\Xset)} \gtrsim \max\{n^{-2s/(4s + d)},n^{-1/4}\}$.
+\textcolor{red}{The same conclusions hold with respect to $\wt{f}_{LS}$.}
+		\item \textbf{Taking advantage of higher-order smoothness conditions.} Under appropriate boundary conditions on the function classes $H^s(\Xset)$ and additional smoothness conditions on the density $p$, when $s > 1$  each of the aforementioned conclusions hold for certain values of $d$.
 		\item \textbf{Manifold adaptivity.}
 		If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$, each of the aforementioned rates hold with $d$ replaced by $m$.
-		\item \textbf{Taking advantage of higher-order smoothness conditions.} Under appropriate boundary conditions on the function classes $C^s$ and $H^s$ and additional smoothness conditions on $P$, when $s > 1$  each of the aforementioned conclusions hold for certain values of $d$.
 	\end{itemize}	
-	\item \textbf{Notation.}
 \end{itemize}
@@
 This restriction is unsurprising, and reflects a fundamental drawback in (penalized) least squares estimators more generally.
-The underlying phenomenon is similar to that elucidated by the seminal papers~\cite{birge1993} and~\cite{birge1998}: when the dimension $d$ is at least $4$, the parameter space is too large, and (penalized) least squares approaches are no longer minimax optimal.
-We should also mention that in a related context (estimation over a $d$-dimensional lattice graph), \cite{sadhanala2016} require the same restriction.
+The underlying phenomenon is similar to that elucidated by the seminal papers~\cite{birge1993} and~\cite{birge1998}: when the dimension $d$ is large enough, the parameter space is too large, and (penalized) least squares approaches are no longer minimax optimal.
+It is interesting to note that we see a gap emerge between the graph- (Laplacian smoothing) and continuum (tensor product spline) estimator when $d = 3$: the former continues to be minimax optimal, whereas the latter no longer is.
+We should also mention that in a related context (estimation over a $d$-dimensional lattice graph), \cite{sadhanala2016} require the same restriction $d \leq 4$.
 \end{itemize}
@@
 \label{sec:higher_order_smoothness_classes}
-The estimation and testing rates derived in Section~\ref{sec:minimax_optimal_graph_Laplacian_methods} exhibit a classic curse of dimensionality: as $d$ gets larger, inference becomes (much) harder.
+The estimation and testing rates derived in Section~\ref{sec:minimax_optimal_graph_Laplacian_methods} reflect a bonafide curse of dimensionality: as  the dimension $d$ gets larger, inference quickly becomes (much) harder.
 As covered in Section~\ref{sec:problem_setup_and_background}, it is well known that by making additional structural assumptions---either on the regression function $f_0$, the distribution $P$, or both---one can derive improved minimax rates.
@@
 \begin{lemma}
-	\label{lem:fixed_graph_testing}
+	\label{lem:le_fixed_graph_testing}
 	Suppose we observe a graph $G$ and responses $Y$ according to the model~\eqref{eqn:fixed_graph_regression_model}. Then for any $1 \leq \kappa \leq n$, each of the following holds for any $b \geq 1$,

### 2020-08-19 notes/weekly_notes/1.30.20.NotesForWeek.tex

@@
 \usepackage[parfill]{parskip}
-\usepackage{algpseudocode}
-\usepackage{algorithm}
 \usepackage{enumerate}
@@
 
-% \newcommand{\span}{\textrm{span}}
-
 \newtheoremstyle{alden}
@@
 
-Let $t$ be an integer, $n = t^d$, and 
+Let $N$ be an integer, $n = N^d$, and 
 \begin{equation*}
-\overline{X} := \Bigl\{\frac{1}{t}(j_1,\ldots,j_d): k \in [t]^d\Bigr\}
+\wb{X} := \Bigl\{\frac{1}{N}(j_1 - 1/2,\ldots,j_d - 1/2): j \in [N]^d\Bigr\}
 \end{equation*}
@@
 \begin{equation*}
-\overline{G} := \bigl(\overline{X},\overline{E}\bigr),~~ \overline{E} := \set{(\overline{x},\overline{x}'): \norm{\overline{x} - \overline{x}'}_1 = \frac{1}{t}}.
+\wb{G} := \bigl(\wb{X},\wb{E}\bigr),~~ \wb{E} := \set{(\wb{x},\wb{x}'): \norm{\wb{x} - \wb{x}'}_1 = \frac{1}{t}}.
 \end{equation*}
@@
 \label{eqn:grid_regression_model}
-y_j = f(\ol{x}_j) + \varepsilon_j, ~\varepsilon_j \overset{\mathrm{i.i.d}}{\sim} \mathcal{N}(0,1) ~~\textrm{for each $j \in [t]^d$,}
+y_j = f(\wb{x}_j) + \varepsilon_j, ~\varepsilon_j \overset{\mathrm{i.i.d}}{\sim} \mathcal{N}(0,1) ~~\textrm{for each $j \in [t]^d$,}
 \end{equation}
@@
 For an arbitrary graph $G$ on $n$ nodes with graph Laplacian $L_G = D_G - A_G$, let $\Lambda(G)$ consist of the eigenvalues $0 = \lambda_1(G) \leq \lambda_2(G) \leq \cdots \leq \lambda_n(G)$ of $L_G$, with $v_k(G) = (v_{k,1}(G),\ldots,v_{k,n}(G)) \in \Reals^n$ denoting the eigenvector corresponding to the $i$th eigenvalues $\lambda_k(G)$.
-The graph Laplacian eigenvector projection test is a truncated-series test; letting $\kappa$ be some integer between $1$ and $n$ which indicates where we truncate the series $\bigl(Y^T v_k(G)\bigr)$, our test statistic is
+The graph Laplacian eigenvector projection test is a linear projection test; letting $\kappa$ be some integer between $1$ and $n$, we define $\Pi_{k,G}(\theta) := \frac{1}{\sqrt{n}}\sum_{k = 1}^{\kappa} \dotp{v_k}{\theta}$ to be a suitable rescaling of the projection of $\theta$ onto the span of $v_1(G),\ldots,v_k(G)$.
+
+Our test statistic
 \begin{equation*}
-T_{\mathrm{spec}}(G) := \frac{1}{n} \sum_{k = 1}^{\kappa} \Biggl(\sum_{i = 1}^{n} y_i v_{k,i}(G)\Biggr)^2
+T_{\mathrm{spec}}(G) := \norm{\Pi_{\kappa,G}(Y)}_n^2 - \frac{\kappa}{n}
 \end{equation*}
-and the test $\phi_{\spec}(G) := \1\{T_{\spec}(G) \geq \tau\}$ rejects the null hypothesis when the test statistic $T_{\mathrm{spec}}(G)$ is greater than a pre-specified cutoff $\tau$.
+is a simple function of the projection of the data $Y$ onto the span of the first $\kappa$ eigenvectors of the graph Laplacian $L(G)$ -- it is a standard calculation to show that
+\begin{equation*}
+\Ebb_{Y|X}\Bigl[\norm{\Pi_{\kappa,G}(Y)}_n^2\Bigr] = \norm{\Pi_{\kappa,G}(f)}_n^2 + \frac{\kappa}{n}
+\end{equation*}
+and hence our test statistic is an unbiased estimate of the norm $\norm{\Pi_{\kappa,G}(f)}_n^2$.
+The graph Laplacian eigenvector test $\phi_{\spec}(G) := \1\{T_{\spec}(G) \geq \tau(b)\}$ rejects the null hypothesis when the test statistic $T_{\mathrm{spec}}(G)$ is greater than a pre-specified cutoff $\tau(b)$, where the number $b$---or more precisely, $1/b$---encodes the level of error the user is willing to tolerate.
 
-The truncation level $\kappa$ and cutoff $\tau$ are user-specified hyperparameters.
-When they are appropriately chosen, the test $\phi_{\spec}(\overline{G})$ is minimax optimal over the Holder ball $C^{1}(\Xset;R)$ when $d \leq 4$.
+The number of eigenvectors $\kappa$ and cutoff $\tau(b)$ are user-specified hyperparameters.
+When they are appropriately chosen, the test $\phi_{\spec}(\wb{G})$ is minimax optimal over the Holder ball $C^{1}(\Xset;R)$ when $d \leq 4$.
 \begin{theorem}
 	\label{thm:holder_testing_rate_grid}
-	Let $b \geq 1$ and $R > 0$ be fixed constants, and let $d \leq 4$.
+	Let $L > 0$ be a fixed constant, and let $d \leq 4$.
 Suppose we observe data according to the model~\eqref{eqn:grid_regression_model}.
-Then when the test $\phi_{\spec}(\overline{G})$ is performed with the parameter choices
+Then when the test $\phi_{\spec}(\wb{G})$ is performed with the parameter choices
 	\begin{equation*}
-	\kappa = n^{2d/(4 + d)},~~ \tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}
+	\kappa = n^{2d/(4 + d)},~~ \tau(b) = b\sqrt{\frac{2\kappa}{n^2}}
 	\end{equation*}
-	the following statement is true: there exists a constant $c$ such that for 
+	for some $b \geq 1$, the following statement is true: there exists a constant $c$ such that for 
 	\begin{equation}
 	\label{eqn:holder_testing_rate_grid_1}
-	\epsilon^2 \geq c R^2 b^2 n^{-4/(4 + d)}
+	\epsilon^2 \geq c L^2 b^2 n^{-4/(4 + d)}
 	\end{equation}
-	the worst-case risk of $\phi_{\spec}(\overline{G})$ on $C^{1}(\Xset;R)$ is upper bounded
+	the worst-case risk of $\phi_{\spec}(\wb{G})$ on $C^{1}(\Xset;L)$ is upper bounded
 	\begin{equation}
 	\label{eqn:holder_testing_rate_grid_2}
-	\mathcal{R}_{\epsilon}\Bigl(\phi_{\spec}(\overline{G});C^{1}(\Xset;R)\Bigr) \leq \frac{1}{b}
+	\mathcal{R}_{\epsilon}\Bigl(\phi_{\spec}(\wb{G});C^{1}(\Xset;L)\Bigr) \leq \frac{1}{b}.
 	\end{equation}
@@
 The test they use to establish the upper bound is related to our own, however, they use a continuum wavelet basis orthogonalized in $L^2([0,1]^d)$ rather than the grid Laplacian eigenvectors we consider.
+
+As was the case with the neighborhood graph in the random design setting, the grid Sobolev norm fails to track the higher order derivatives of $f$ near the boundary of $\Xset$.
+We therefore must insist that $f$ be compactly supported in order for the grid eigenvector projection test to achieve minimax optimal testing rates over higher-order Holder classes.
+
+\begin{theorem}
+	\label{thm:holder_testing_rate_grid_higher_order}
+	Let $L > 0$ be fixed constants, $s \geq 2$ be a fixed integer, and let $d \leq 4s$.
+Suppose we observe data according to the model~\eqref{eqn:grid_regression_model}.
+Then when the test $\phi_{\spec}(\wb{G})$ is performed with the parameter choices
+	\begin{equation*}
+	\kappa = n^{2d/(4s + d)},~~ \tau(b) =  b\sqrt{\frac{2\kappa}{n^2}}
+	\end{equation*} 
+	for some $b \geq 1$, the following statement is true: there exists a constant $c$ such that for any
+	\begin{equation*}
+	\epsilon^2 \geq c L^2 b^2 n^{-4s/(4s + d)},
+	\end{equation*}
+	the worst-case risk of $\phi_{\spec}(\wb{G})$ is upper bounded
+	\begin{equation*}
+	\mathcal{R}_{\epsilon}\Bigl(\phi_{\spec}(\wb{G});C_c^{s}(\Xset;L)\Bigr) \leq \frac{1}{b}.
+	\end{equation*}
+\end{theorem}
+
+In order to prove Theorems~\ref{thm:holder_testing_rate_grid} and~\ref{thm:holder_testing_rate_grid_higher_order}, we decompose
+\begin{equation*}
+T_{\spec}(\wb{G}) - \norm{f}_{\Leb^2}^2 = \Bigl\{\norm{\Pi_{\kappa,\wb{G}}Y}_n^2 - \frac{\kappa}{n} - \norm{\Pi_{\kappa,\wb{G}}f}_n^2\Bigr\} + \Bigl\{\norm{\Pi_{\kappa,\wb{G}}f}_n^2 - \norm{f}_n^2\Bigr\} + \Bigl\{\norm{f}_n^2 - \norm{f}_{\Leb^2}^2\Bigr\},
+\end{equation*}
+where we refer to the three terms on the right hand side as the estimation error, approximation error, and the discretization error, respectively.
+
+Our technical work consists of upper bounding these three sources of error.
+When $4s \geq d$, the discretization error is negligible regardless of our choice of $\kappa$, and we focus our attention on minimizing the sum of approximation and estimation errors.
+This is the standard tradeoff in nonparametric testing problems, and the usual choice of projecting our data $Y \in \Reals^n$ onto a subspace of dimension $\kappa = n^{2d/(4s + d)}$ balances the contribution of these two terms.
+
+\subsection{Low-smoothness regime}
+In the low-smoothness regime $d > 4s$, the coupling between empirical norm over $\wb{X}$ and the continuum $\Leb^2$ norm is sufficiently weak that the discretization error can become the dominant source of error, as we show in Lemma~\ref{lem:holder_testing_rate_grid_low_smoothness_lb}.
+
+\begin{lemma}
+	\label{lem:holder_testing_rate_grid_low_smoothness_lb}
+	For any $L > 0$ and any integers $s$ and $d$, there exists a function $f \in C^s(\Xset;L)$ such that
+	\begin{equation*}
+	\norm{f}_{\Leb^2}^2 \geq \frac{1}{2^d} L^2 n^{-2s/d} 
+	\end{equation*}
+	but $f(\wb{x}) = 0$ for all $\wb{x} \in \wb{X}$. 
+\end{lemma}
+Lemma~\ref{lem:holder_testing_rate_grid_low_smoothness_lb} implies the critical radius over the Holder classes $C^s(\Xset;L)$ must be at least on the order of $\epsilon^{\star}(C^s(\Xset;L)) \gtrsim n^{-s/d}$.
+When $4s \geq d$, this term is neglible relative to $n^{-2s/(4s + d)}$, and we arrive at the ``usual'' rates of minimax testing we see in Theorems~\ref{thm:holder_testing_rate_grid} and~\ref{thm:holder_testing_rate_grid_higher_order}.
+On the other hand when $4s < d$, this becomes the dominant term, and we can no longer achieve the typical rate $\epsilon^2 \asymp n^{-4s/(4s + d)}$.
+
+In this setting, when we enforce the wider radius $\epsilon^2 \asymp n^{-2s/d}$, Proposition~\ref{prop:holder_testing_rate_grid_low_smoothness_ub} shows that a very simply test has small worst-case risk $\mathcal{R}_{\epsilon}$.
+
+\begin{proposition}
+	\label{prop:holder_testing_rate_grid_low_smoothness_ub}
+	Let $L > 0$ be a fixed constant, $s \geq 1$ a fixed integer, and let $d \geq 4s$.
+Suppose we observe data according to the grid design regression model~\eqref{eqn:grid_regression_model}.
+Then when the test $\phi_{\spec}(\wb{G})$ is performed with the parameter choices
+	\begin{equation*}
+	\kappa = n,~~ \tau(b) = b\sqrt{\frac{2}{n}}
+	\end{equation*}
+	for some $b \geq 1$, the following statement is true: there exists a constant $c$ such that for any 
+	\begin{equation*}
+	\epsilon^2 \geq c b^2 L^2 n^{-2s/d}
+	\end{equation*}
+	the worst case risk of $\phi_{\spec}(\wb{G})$ is upper bounded
+	\begin{equation*}
+	\mathcal{R}_{\epsilon}(\phi_{\spec}(\wb{G}); C^s(\Xset;L)) \leq \frac{1}{b}. 
+	\end{equation*}
+\end{proposition}
+
+\begin{itemize}
+	\item The test statistic $T_{\spec}(\wb{G})$ is simply the empirical norm of $Y$. 
+	\item Together, Lemma~\ref{lem:holder_testing_rate_grid_low_smoothness_lb} and Proposition~\eqref{prop:holder_testing_rate_grid_low_smoothness_ub} characterize the minimax rate of the grid design regression testing problem when $4s > d$.
+	\item This minimax rate does not match the upper bound established in Proposition~\textcolor{red}{(Holder testing random design)}---the fixed design problem is harder. 
+	\item Note that the usual minimax rate of estimation error over Holder classes is always $n^{-2s/(2s + d)}$, regardless of the relation between $s$ and $d$, and regardless of whether loss is measured in $\Leb^2$-norm or empirical norm.
+The explanation for this is that the rate $n^{-2s/(2s + d)}$ is always larger than $n^{-2s/d}$, and so the discretization error is never the dominant source of error in estimation.
+This reveals an interesting distinction between the testing and estimation problems, as testing is easy enough in a statistical sense that error incurred by the discrete nature of the problem may be the bottleneck.
+\end{itemize}
+
+\subsection{Continuum Sobolev class.}
+
+Now suppose we assume $f \in H^s(\Xset;L)$ belongs to a relevant Sobolev ball, rather than a Holder ball, and we again ask about the worst case risk $\mathcal{R}_{\epsilon}(\phi;H^s(\Xset;L))$.
+
+When $2s < d$, the situation is hopeless.
+The Sobolev class $H^s(\Xset;L)$ includes bump functions of arbitrarily small radius and arbitrarily large height---in the limit, delta functions.
+By placing these bump functions at design points $\wb{x}$, we can construct a function $f$ which satisfies $f \equiv L$ almost everywhere on $\Xset$, while $f(\wb{x}) = 0$ for all $\wb{x} \in \wb{X}$.
+No test based on data observed according to model~\eqref{eqn:grid_regression_model} can have more than trivial power against $f$, implying that for all tests $\phi$ the worst case risk $\mathcal{R}_{\epsilon}(\phi;H^s(\Xset;L))$ is $1$ unless $\epsilon = \Omega(1)$.
+
+When $2s > d$, we are not sure how the situation changes. The Sobolev Embedding Theorem gives us that the critical radius 
+\begin{equation*}
+\epsilon^{\star}\bigl(H^s(\Xset;L)\bigr) \leq c \epsilon^{\star}\bigl(C^{s'}(\Xset;L)\bigr),~~\textrm{for $s' = s - d/2$.}
+\end{equation*}
+where we recall that the critical radius over Holder functions is
+\begin{equation*}
+\epsilon^{\star}\bigl(C^{s}(\Xset;L)\bigr) \asymp \max\bigl\{n^{-2s/(4s + d)}, n^{-s/d}\bigr\}
+\end{equation*}
+We might hope to do better, but it is hard to say.
 
@@
 	\begin{equation*}
-	\tau = \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}.
+	\tau = b\sqrt{\frac{2\kappa}{n^2}}.
 	\end{equation*}
@@
 The proof of Lemma~\ref{lem:fixed_graph_testing} can be seen in the \textit{graph\_testing} document.
-The eigenvalues of the lattice $\ol{G}$ are known exactly, and standard manipulations \textcolor{red}{Sadhanala} lead to the lower bound
+The eigenvalues of the lattice $\wb{G}$ are known exactly, and standard manipulations \textcolor{red}{Sadhanala} lead to the lower bound
 \begin{equation}
 \label{eqn:eigenvalue_tail_bound_grid}
-\lambda_k(\ol{G}) \geq c n^{-2/d}k^{2/d}~\textrm{for $k = 1,\ldots,n$.}
+\lambda_k(\wb{G}) \geq c n^{-2/d}k^{2/d}~\textrm{for $k = 1,\ldots,n$.}
 \end{equation}
@@
 	\label{eqn:discrete_sobolev_norm_grid}
-	f^T \bigl(L_{\ol{G}}\bigr) f \leq 4R^2 n^{1 - 2/d}
+	f^T \bigl(L_{\wb{G}}\bigr) f \leq 4R^2 n^{1 - 2/d}
 	\end{equation}
@@
 \begin{align*}
-2b\sqrt{\frac{2\kappa}{n^2}} + \frac{f^T \bigl(L_{\ol{G}}\bigr)^s f}{n\bigl[\lambda_{\kappa}(\ol{G})\bigr]^s} \leq \frac{1}{n}\sum_{i = 1}^{n}\bigl[f(x_i)\bigr]^2
+2b\sqrt{\frac{2\kappa}{n^2}} + \frac{f^T \bigl(L_{\wb{G}}\bigr)^s f}{n\bigl[\lambda_{\kappa}(\wb{G})\bigr]^s} \leq \frac{1}{n}\sum_{i = 1}^{n}\bigl[f(x_i)\bigr]^2
 \end{align*}
@@
 \begin{align*}
-f^T (L_{\ol{G}}) f & = \sum_{i \in [t]^d} \sum_{j \in [t]^d} \bigl(f(\ol{x}_i) - f(\ol{x}_j\bigr)^2 \1\{\norm{i - j}_1 = t\} \\
-& \leq R^2 \sum_{i \in [t]^d} \sum_{j \in [t]^d} \norm{\ol{x}_i - \ol{x}_j}_2^2 \cdot  \1\{\norm{i - j}_1 = t\} \\
+f^T (L_{\wb{G}}) f & = \sum_{i \in [t]^d} \sum_{j \in [t]^d} \bigl(f(\wb{x}_i) - f(\wb{x}_j\bigr)^2 \1\{\norm{i - j}_1 = t\} \\
+& \leq R^2 \sum_{i \in [t]^d} \sum_{j \in [t]^d} \norm{\wb{x}_i - \wb{x}_j}_2^2 \cdot  \1\{\norm{i - j}_1 = t\} \\
 & = R^2 n^{-2/d} \sum_{i \in [t]^d} \sum_{j \in [t]^d} \1\{\norm{i - j}_1 = t\}  \\
@@
 \end{equation*}
-The difference between $f(\ol{x})$ and the average of $f$ over the cube $Q(\ol{x})$ can be bounded using the Holder property,
+The difference between $f(\wb{x})$ and the average of $f$ over the cube $Q(\wb{x})$ can be bounded using the Holder property,
 \begin{equation*}
-\int_{Q(\ol{x})} (f(x) - f(\ol{x}))^2 \,dx \leq \frac{1}{n} R^2 n^{-2/d}.
+\int_{Q(\wb{x})} (f(x) - f(\wb{x}))^2 \,dx \leq \frac{1}{n} R^2 n^{-2/d}.
 \end{equation*}
@@
 \begin{align*}
-\norm{f}_{\Leb^2(\Xset)}^2 & = \sum_{j \in [t]^d} \int_{Q(\ol{x}_j)} \bigl[f(x)\bigr]^2 \,dx \\
-& \leq 2 \sum_{j \in [t]^d} \frac{1}{n} \Bigl\{\bigl[f(\ol{x}_j)\bigr]^2 + \int_{Q(\ol{x}_j)} \bigl[f(x) - f(\ol{x}_j)]^2 \,dx \Bigr\} \\
-& \leq \frac{2}{n}\sum_{j \in [t]^d} \bigl[f(\ol{x}_j)\bigr]^2 + R^2n^{-2/d} \\
-& \leq \frac{2}{n}\sum_{j \in [t]^d} \bigl[f(\ol{x}_j)\bigr]^2 + \frac{1}{2}\norm{f}_{\Leb^2(\Xset)}^2
+\norm{f}_{\Leb^2(\Xset)}^2 & = \sum_{j \in [t]^d} \int_{Q(\wb{x}_j)} \bigl[f(x)\bigr]^2 \,dx \\
+& \leq 2 \sum_{j \in [t]^d} \frac{1}{n} \Bigl\{\bigl[f(\wb{x}_j)\bigr]^2 + \int_{Q(\wb{x}_j)} \bigl[f(x) - f(\wb{x}_j)]^2 \,dx \Bigr\} \\
+& \leq \frac{2}{n}\sum_{j \in [t]^d} \bigl[f(\wb{x}_j)\bigr]^2 + R^2n^{-2/d} \\
+& \leq \frac{2}{n}\sum_{j \in [t]^d} \bigl[f(\wb{x}_j)\bigr]^2 + \frac{1}{2}\norm{f}_{\Leb^2(\Xset)}^2
 \end{align*}
@@
 
+\subsection{Proof of Theorem~\ref{thm:holder_testing_rate_grid_higher_order}}
+Referring again to Lemma~\ref{lem:fixed_graph_testing}, to prove Theorem~\ref{thm:holder_testing_rate_grid_higher_order} it will be sufficient to prove the following inequalities
+\begin{enumerate}[label=(E\arabic*)]
+	\item 
+	\label{event:grid_sobolev_norm_higher_order}
+	For any $f \in C_c^s(\Xset;L)$,
+	\begin{equation}
+	\label{eqn:grid_sobolev_norm_higher_order}
+	f^T \wb{L}^s f \leq c L^2 n^{1 - 2s/d}.
+	\end{equation}
+	\item 
+	\label{event:grid_empirical_norm_higher_order}
+	There exists a sufficiently small constant $c_1$ and a sufficiently large constant $c_2$ such that for any $f \in C^s(\Xset;L)$,
+	\begin{equation}
+	\label{eqn:grid_empirical_norm_higher_order}
+	c_1 \norm{f}_n^2 \geq \norm{f}_{\Leb^2(\Xset)}^2 - c_2 L n^{-2s/d}
+	\end{equation}
+	for all $n$ sufficiently large.
+\end{enumerate}
+The proof of \eqref{eqn:grid_empirical_norm_higher_order} is the content of Lemma~\ref{lem:discretization_error_holder}.
+Here, we focus our attention on proving~\eqref{eqn:grid_sobolev_norm_higher_order}.
+
+To obtain this upper bound, first observe that by letting $D_{e_i^2}f := f(\wb{x} + he_i) + f(\wb{x} - he_i) - 2f(\wb{x})$ and taking a $2$nd order Taylor expansion, for all points $\wb{x} \in \Xset_{h} \cap \wb{X} =: \wb{X}_{h}$ we can express the grid Laplacian as
+\begin{align}
+\wb{L}f(\wb{x}) & = \sum_{i = 1}^{d} D_{e_i^2}f(\wb{x}) \nonumber \\
+& = \sum_{i = 1}^{d} \biggl[\frac{\partial}{\partial e_i} f(\wb{x})(e_ih - e_ih) + \frac{h^2}{2} \int_{-h}^{h} \frac{\partial^2}{\partial e_i^2} f(\wb{x} + e_it) \,dt \biggr] \nonumber \\
+& = \frac{h^2}{2} \int_{-h}^{h} \sum_{i = 1}^{d} \frac{\partial^2}{\partial e_i^2} f(\wb{x} + e_it) \,dt \label{eqn:grid_sobolev_norm_higher_order_pf1}
+\end{align} 
+as a second-order differential operator which roughly approximates the continuum Laplace operator $\Delta = -\sum_{i = 1}^{d} \frac{\partial^2}{\partial e_i^2}$.
+Motivated by this, in Lemma~\ref{lem:grid_laplacian_approximation_error} we provide estimates on the absolute difference between the action of the iterated grid Laplacian and the iterated continuum Laplace operator.
+Specifically, we show that at all sufficiently interior points $\wb{x} \in \wb{X} \cap \Xset_{qh}$
+\begin{equation*}
+\begin{rcases*}
+\abs{\wb{L}^{(s - 2)/2}f(\wb{x}) - h^{(s - 2)/2}\Delta^{(s - 2)/2}f(\wb{x})}, &\textrm{when $s$ is even} \\
+\abs{\wb{L}^{(s - 1)/2}f(\wb{x}) - h^{(s - 1)/2}\Delta^{(s - 1)/2}f(\wb{x})}, & \textrm{when $s$ is odd} \\
+\end{rcases*}
+\leq cLh^{s}
+\end{equation*}
+For those remaining points $\wb{x} \in \wb{X} \cap \Xset_{qh}$ sufficiently near the boundary of $\Xset$, since $f$ is compactly supported in $\Xset$ there exists a point $x_0$ and some $\delta > 0$ such that $\mathrm{dist}(\wb{x},x_0) < qh$, and $f(z) = 0$ for all $z \in B(x_0,r)$.
+Therefore the partial derivatives $D^qf(x_0) = 0$ for each $q = 0,\ldots,s - 1$, and by taking a Taylor expansion of $f(x)$ around $f(x_0)$ we obtain
+\begin{equation*}
+\abs{f(x)} \leq c L h^{s}
+\end{equation*}
+where we have additionally used the Holder property of $f$. A crude bound of the form
+\begin{equation*}
+\abs{\wb{L}^s f(\wb{x})} \leq 2^s \cdot \max_{\wb{z} \in Q_{2qh}(\wb{x})} \bigl\{\abs{f(\wb{z})} \bigr\} \leq c L h^{s}
+\end{equation*}
+will suffice for our purposes. When $s$ is even, the graph Sobolev seminorm is then
+\begin{align*}
+f^T \wb{L}^s f & \leq \sum_{\wb{x} \in \wb{X}_{qh}} \Bigl\{\sum_{i = 1}^d D_{e_i^2} \wb{L}^{(s - 2)/2}f(\wb{x}) \Bigr\}^2 + cL^2h^{2s}n^{1 - 1/d} \\
+& \leq h^{2s - 4}\sum_{\wb{x} \in \wb{X}_{qh}} \Bigl\{\sum_{i = 1}^d D_{e_i^2} \Delta^{(s - 2)/2}f(\wb{x}) \Bigr\}^2 + cL^2h^{2s}n^{1 - 1/d} \\
+& \leq \frac{h^{2s}}{4}\sum_{\wb{x} \in \wb{X}_{qh}} \Bigl\{\sum_{i = 1}^d \int_{-h}^{h} \Bigl(\frac{\partial^2}{\partial e_i^2}\Delta^{(s - 2)/2}f\Bigr)(\wb{x} + te_i) \,dt \Bigr\}^2 + cL^2h^{2s}n^{1 - 1/d} \\
+& \leq d^2 h^{2s} L^2 n + cL^2h^{2s}n^{1 - 1/d}
+\end{align*}
+where we have used, in order, our bounds on the absolute value of $\wb{L}^sf(\wb{x})$ and $\abs{f}(\wb{x})$ at points $\wb{x}$ near the boundary, our estimate on the approximation error between the grid Laplacian and continuum Laplace operator,
+the representation~\eqref{eqn:grid_sobolev_norm_higher_order_pf1}, and finally the Holder property
+\begin{equation*}
+\norm{\Delta^{(s - 2)/2}f}_{C^2(\Xset)} \leq \norm{f}_{C^s(\Xset)} \leq L.
+\end{equation*}
+Similar manipulations supply an equivalent bound when $s$ is odd
+\begin{align*}
+f^T \wb{L}^s f & \leq \sum_{\wb{x} \in \wb{X}_{qh}} \Bigl\{\sum_{i = 1}^d \wb{L}^{(s - 1)/2}f(\wb{x} + he_i) - \wb{L}^{(s - 1)/2}f(\wb{x}) \Bigr\}^2 + cL^2h^{2s}n^{1 - 1/d} \\
+& \leq h^{2s - 2}\sum_{\wb{x} \in \wb{X}_{qh}} \Bigl\{\sum_{i = 1}^d \wb{\Delta}^{(s - 1)/2}f(\wb{x} + he_i) - \wb{\Delta}^{(s - 1)/2}f(\wb{x}) \Bigr\}^2 + cL^2h^{2s}n^{1 - 1/d} \\
+& \leq d^2 h^{2s} L^2 n + cL^2h^{2s}n^{1 - 1/d}
+\end{align*}
+where this time we have invoked the Holder property $\norm{\Delta^{(s - 1)/2}f}_{C^1(\Xset)} \leq \norm{f}_{C^s(\Xset)} \leq L$.
+This establishes~\eqref{eqn:grid_sobolev_norm_higher_order} and concludes the proof of Theorem~\ref{thm:holder_testing_rate_grid_higher_order}.
+
+\subsection{Proof of Lemma~\ref{lem:holder_testing_rate_grid_low_smoothness_lb}}
+
+\textcolor{red}{TODO:}
+
+\subsection{Proof of Proposition~\ref{prop:holder_testing_rate_grid_low_smoothness_ub}}
+
+\textcolor{red}{TODO:}
+
+\section{Proofs of Technical Results}
+
+\subsection{Grid Laplacian Approximation Error}
+
+The following result follows by standard Taylor expansion arguments.
+
+\begin{lemma}
+	\label{lem:grid_laplacian_approximation_error}
+	Let $f \in C^s(\Xset;L)$ for $s \geq 3$. Then, for all grid points $\wb{x} \in \wb{X}_{qh}$, and for either $q = (s - 1)/2$ or $q = (s - 2)/2$,
+	\begin{equation*}
+	\abs{\wb{L}^{q}f(\wb{x}) - h^{2q}\Delta^{q}f(\wb{x})} \leq cLh^s.
+	\end{equation*}
+\end{lemma}
+\begin{proof}
+	We prove by induction on $q$. When $q = 1$ and $\wb{x} \in \wb{X}_{h}$, by taking Taylor expansions of $f(\wb{x} + e_ih)$ and $f(\wb{x} - e_ih)$ around $f(\wb{x})$ for $i = 1,\ldots,d$, we obtain
+	\begin{align}
+	\wb{L}f(\wb{x}) & = \sum_{i = 1}^{d} D_{e_i^2}f(\wb{x}) \nonumber \\
+	& = \sum_{i = 1}^{d} \Biggl[\sum_{\alpha = 1}^{s - 1}\frac{1}{\alpha!}\frac{\partial^{\alpha}}{\partial e_i^{\alpha}} f(\wb{x})\bigl(h^{\alpha} + (-h)^{\alpha}\bigr) + R(\wb{x}) \Biggr] \nonumber \\
+	& = \sum_{i = 1}^{d} \Biggl[\sum_{\beta = 1}^{\floor{(s - 1)/2}}\frac{2}{(2\beta)!}\frac{\partial^{2\beta}}{\partial e_i^{2\beta}} f(\wb{x}) h^{2\beta} + R(\wb{x})\Biggr] \label{eqn:grid_laplacian_approximation_error_pf1}
+	\end{align} 
+	where the remainder term is upper bounded
+	\begin{equation*}
+	\abs{R(\wb{x})} = \abs{\frac{1}{s!}\sum_{i = 1}^d \int_{-h}^{h} \frac{\partial^s}{\partial e_i^s} f(\wb{x} + e_it) t^{s - 1} \,dt} \leq \frac{d}{(s + 1)!} L h^s 
+	\end{equation*}
+	When $s = 3$ or $s = 4$, the first sum includes only second-order derivatives,
+	\begin{equation*}
+	\sum_{i = 1}^{d} \sum_{\beta = 1}^{\floor{(s - 1)/2}}\frac{2}{(2\beta)!}\frac{\partial^{2\beta}}{\partial e_i^{2\beta}} f(\wb{x}) h^{2\beta} = h^2 \sum_{i = 1}^d \frac{\partial^{2\beta}}{\partial e_i^{2\beta}} f(\wb{x}) = h^2 \Delta f(\wb{x}),
+	\end{equation*}
+	and the claim is shown.
+
+	When $q \geq 2$, we use the prior analysis to rewrite the $q$th-order iterated Laplacian
+	\begin{align}
+	\wb{L}^qf(\wb{x}) & = \wb{L}^{q - 1} \Biggl(\sum_{i = 1}^d \sum_{\beta = 1}^{\floor{(s - 1)/2}} \frac{2}{(2\beta)!} \frac{\partial^{2\beta}}{\partial e_i^{2\beta}} f h^{2 \beta} + R\Biggr)(\wb{x}) \nonumber \\
+	& = \sum_{\beta = 1}^{\floor{(s - 1)/2}} \frac{2h^{2\beta}}{(2\beta)!} \wb{L}^{q - 1} \biggl(\sum_{i = 1}^d\frac{\partial^{2\beta}}{\partial e_i^{2\beta}} f \biggr)(\wb{x}) + \wb{L}^{q - 1}R(\wb{x}) \label{eqn:grid_laplacian_approximation_error_pf5}
+	\end{align}
+	We deal with the remainder term first. Noting that $Q_{2(q - 1)h}(\wb{x}) \subset \Xset_{h}$, we have
+	\begin{align}
+	\abs{\wb{L}^{q - 1}R(\wb{x})} & \leq \bigl(2(q - 1)d\bigr)^{q - 1} \cdot \max_{\wb{z} \in Q_{2(q - 1)h}(\wb{x})} \bigl\{\abs{R(\wb{z})}\bigr\} \nonumber \\
+	& \leq \frac{\bigl(2(q - 1)d\bigr)^{q - 1}d}{(s + 1)!} L h^s \label{eqn:grid_laplacian_approximation_error_pf2}
+	\end{align}
+	Next we deal with the higher-order terms in the first sum, meaning the terms where $\beta \geq 2$.
+For convenience, denote $f_\beta := \sum_{i = 1}^d\frac{\partial^{2\beta}}{\partial e_i^{2\beta}} f$, and note that $f_{\beta} \in C^{s - 2\beta}(\Xset;L)$.
+
+	We will show that for any $\wb{x} \in \wb{X}_{qh}$,
+	\begin{equation}
+	\label{eqn:grid_laplacian_approximation_error_pf3}
+	\abs{\wb{L}^{q - 1}f_{\beta}(\wb{x})} \leq c L h^{2 - \beta}.
+	\end{equation}
+	Let us assume $s - 2\beta$ is odd. Set $s' = (s - 2\beta - 1)/2$, and $s'' = q - 1 - (s' + 1)$, and note that $s''$ is a non-negative integer. By hypothesis, for any $\wb{z} \in \wb{X}_{hs'}$
+	\begin{equation*}
+	\abs{\wb{L}^{s'}f_{\beta}(\wb{z}) - h^{2s'}\Delta^{s'}f_{\beta}(\wb{z})} \leq cLh^{s - 2\beta}
+	\end{equation*}
+	and consequently,
+	\begin{align*}
+	\abs{\wb{L}^{q - 1}f_{\beta}(\wb{x})} & \leq (2s''d)^{s''} \cdot \max_{\wb{z} \in Q_{2s''h}(\wb{x})} \Bigl\{\abs{\wb{L}^{s' + 1}f(\wb{z})}\Bigr\} \\
+	& \leq (2s''d)^{s''} \cdot \max_{\wb{z} \in \wb{X}_{(s' + 1)h}} \Bigl\{\abs{\wb{L}^{s' + 1}f(\wb{z})}\Bigr\}  \\
+	& \leq (2s''d)^{s''} \Bigl( h^{2s} \cdot \max_{\wb{z} \in \wb{X}_{(s' + 1)h}} \Bigl\{ \abs{\wb{L} \Delta^{s'} f_{\beta}(\wb{z})} \Bigr\} + cLh^{s - 2\beta}\Bigr) \\
+	& \leq (2s''d)^{s''} \Bigl( 2dL h^{2s' + 1}  + cLh^{s - 2\beta}\Bigr)
+	\end{align*}
+	where the last line follows since $\Delta^{s'}{f_{\beta}} \in C^1(\ol{X};L)$.
+Since $2s' + 1 = s - 2\beta$, we have shown the desired result~\eqref{eqn:grid_laplacian_approximation_error_pf3} when $s - 2\beta$ is odd, and similar manipulations prove the result when $s - 2\beta$ is even.
+
+	Finally, we consider the second order term, where $\beta = 2$ and $f_\beta = \Delta f \in C^{s - 2}(\Xset;L)$.
+Then, since either $q - 1 = (s - 3)/2$ or $q - 1 = (s - 4)/2$, by hypothesis for any $\wb{x} \in \wb{X}_{qh} \subset \wb{X}_{(q - 1)h} $,
+	\begin{equation}
+	\label{eqn:grid_laplacian_approximation_error_pf4}
+	h^{2}\abs{ \wb{L}^{q - 1} \Delta f(\wb{x}) - h^{2(q - 1)} \Delta^{q} f(\wb{x})  } \leq cL h^{2}h^{s - 2} = cLh^s
+	\end{equation}
+The claimed result follows upon plugging~\eqref{eqn:grid_laplacian_approximation_error_pf2},\eqref{eqn:grid_laplacian_approximation_error_pf3}, and \eqref{eqn:grid_laplacian_approximation_error_pf4} back into~\eqref{eqn:grid_laplacian_approximation_error_pf5}.
+\end{proof}
+
+\subsection{Discretization Error}
+We prove the following lower bound on the empirical norm of $f$.
+The proof will be similar in spirit and at times directly follow the proof of Lemma 3 of \textcolor{red}{(Arias-Castro)}, the key difference being that we wish to lower bound the empirical norm of $f$ rather than a Riemann sum.
+\begin{lemma}
+	\label{lem:discretization_error_holder}
+	Let $f \in C^s(\Xset,L)$. Then, there exist a constant $c$ such that
+	\begin{equation*}
+	\norm{f}_{n}^2 \geq c\norm{f}_{\Leb^2([0,1]^d)}^2 - c L n^{-2s/d}
+	\end{equation*}
+	for all $n$ sufficiently large.
+\end{lemma}
+\begin{proof}
+	In the proof of this Lemma, we take $\eta$ to be the tensorization of a univariate $(s - 1)$-order kernel $\nu$. More specifically, we let $\nu: \Reals \to \Reals$ satisfy
+	\begin{equation*}
+	\int_{\Reals} \nu(x) \,dx = 1,~~ \int x^{\ell} \nu(x) \,dx = 0~~\textrm{for $\ell = 1,\ldots,s-1$},~~ \abs{\nu(x)} \leq 1~\textrm{for all $x \in [0,1^d]$}
+	\end{equation*}
+	additionally be compactly supported on $[-1/2,1/2]$ but otherwise arbitrary, and define $\eta:\Reals^d \to \Reals$ to be $\eta(z) = \prod_{i = 1}^{d} \nu(x_i)$ for $z \in \Reals^d$.
+We denote $\eta_h(x) = \frac{1}{h^d} \eta(x/h)$, and note that for any polynomial $u$ of degree at most $s - 1$, $\bigl(u \ast \eta_h\bigr)(x) = u(x)$.
+
+	Fix an odd integer $m \geq 1$ to be chosen large enough, but fixed, later on, and assume without loss of generality that $N/m =: M \in \mathbb{N}$.
+For $i \in [n]^d$, let $u_{i}$ be the $(s - 1)$-order Taylor expansion of $f$ around $\wb{x}_i$, and further let
+	\begin{align*}
+	u(x) = \sum_{i \in [M]^d} u_{mi - 1}(x) \cdot \1\{x \in Q_{hm}(\wb{x}_{mi - 1})  \},~~ \wt{u}(x) = \sum_{i \in [M]^d} \sum_{\wb{x} \in  Q_{hm}(\wb{x}_{mi - 1})} \bigl(u_{mi - 1} \ast \eta_h\bigr)(\wb{x}) \cdot \1\{x \in Q_{h}(\wb{x})  \}.
+	\end{align*}
+	be piecewise polynomial and piecewise constant functions, respectively.
+
+	Standard properties of Taylor expansions imply that $\abs{f(x) - u_{mi - 1}(\wb{x})} \leq (L\sqrt{d}mh)^{s}$ for all $x \in Q_{mh}(\wb{x}_{mi - 1})$, and as a result
+	\begin{equation*}
+	\norm{u - f}_{\Leb^2([0,1]^d)}^2,~~\norm{u - f}_{n}^2 \leq ch^{2s}
+	\end{equation*}
+	Since $u_{mi- 1}$ is an $(s - 1)$-order polynomial, we have that $\bigl(u_{mi - 1} \ast \eta_h\bigr)(\wb{x}) = u_{mi - 1}(\wb{x})$ for each $\wb{x} \in \wb{X}$; therefore $\norm{\wt{u}}_n = \norm{u}_n$.
+	Additionally $\wt{u}$ is piecewise constant over the grid cells $Q_h(\wb{x})$, and so we have
+	\begin{equation*}
+	\norm{\wt{u}}_{\Leb^2([0,1]^d)}^2 = \sum_{\wb{x} \in \wb{X}} h^d \cdot  [\wt{u}(\wb{x})]^2 = \norm{\wt{u}}_n^2.
+	\end{equation*}
+	It remains only to show that
+	\begin{equation*}
+	c \norm{u}_{\Leb^2([0,1]^d)} \leq \norm{\wt{u}}_{\Leb^2([0,1]^d)},
+	\end{equation*}
+	for a constant $c$ which does not depend on $\wt{u}$.
+In Lemma~\ref{lem:riemann_sums_polynomials}, we show that the statement holds true if $u$ is a polynomial on $[0,1]^d$; and the width of the partition used to form $\wt{u}$ is taken to be a sufficiently small constant.
+By translating and rescaling the cubes $Q_{mi - 1}(\wb{x}_{mi} - 1)$ into $[0,1]^d$ as in \textcolor{red}{(Arias-Castro)}, we have that the same result holds for the piecewise polynomial $u$ when $m$ is taken to be a sufficiently large constant.
+Putting the pieces together, we have that whenever $n$ is great enough such that $M \in \mathbb{N}$,
+	\begin{align*}
+	\norm{f}_{\Leb^2([0,1]^d)}^2 & \leq cn^{-2s/d} + 2\norm{u}_{\Leb^2([0,1]^d)}^2 \\ & \leq cn^{-2s/d} + c\norm{\wt{u}}_{\Leb^2([0,1]^d)}^2 \\
+	& = cn^{-2s/d} + c\norm{u}_{n}^2 \\
+	& \leq cn^{-2s/d} + c\norm{f}_n^2.
+	\end{align*}
+	which completes the proof of Lemma~\ref{lem:discretization_error_holder}.
+\end{proof}
+
+To simplify the statement of the following Lemma, for $j \in \mathbb{N}$ let
+\begin{equation*}
+\mathcal{K}_j(u) := \sum_{[j]^d} \bigl(u \ast \eta_h\bigr)(\wb{x}) \cdot \1\{x \in Q_{1/j}(\wb{x})\}.
+\end{equation*}
+where $\eta_h$ is taken as in the proof of Lemma~\ref{lem:discretization_error_holder}. Denote the class of polynomials on $\Reals^d$ of degree at most $m$ by $\mathcal{P}_m^d$.
+
+\begin{lemma}
+	\label{lem:riemann_sums_polynomials}
+	There exists a constant $c$ such that for any $j > 1/c$, 
+	\begin{equation*}
+	\norm{\mathcal{K}_j(u)}_{\Leb^2([0,1]^d)} \geq c \norm{u}_{\Leb^2([0,1]^d)}
+	\end{equation*}
+	for all polynomials $u \in \mathcal{P}_{s - 1}^d$.
+\end{lemma}
+\begin{proof}
+	We proceed by contradiction, following the proof of~\textcolor{red}{(Arias-Castro)} closely and making necessary adjustments.
+Suppose there exists a sequence $(u_j) \subset \mathcal{P}_{s - 1}^d$ such that $\norm{\mathcal{K}_j(u)}_{\Leb^2([0,1]^d)} \leq 1/j \norm{u_j}_{\Leb^2([0,1]^d)}$ for each $j \in \mathcal{N}$.
+Assume without loss of generality that $\norm{u_j}_{\Leb^2([0,1]^d)} = 1$; then there exists an accumulation point $u_{\infty}$ of $(u_j)$, and $\norm{u_{\infty}}_{\Leb^2([0,1]^d)} = 1$.
+However, by Jensen's inequality, and the boundedness and compact support of $\eta$, we have
+	\begin{align*}
+	\Bigl[\bigl(u \ast \eta_h\bigr)(\wb{x})\Bigr]^2 & = \Bigl[\int u(y) \eta_h(y - \wb{x}) \,dy \Bigr]^2 \\
+	& = \biggl[\frac{1}{h^d}\int u(y) \eta\Bigl(\frac{y - \wb{x}}{h}\Bigr) \,dy \biggr]^2 \\
+	& \leq \frac{1}{h^d}  \biggl[u(y) \eta\Bigl(\frac{y - \wb{x}}{h}\Bigr)\biggr]^2 \,dy \\
+	& \leq \frac{1}{h^d} \int_{Q_h(\wb{x})} \bigl[u(y)\bigr]^2 \,dy
+	\end{align*}
+	which in turn gives 
+	\begin{align*}
+	\norm{\mathcal{K}_j(u)}_{\Leb^2([0,1]^d)}^2 & \leq \sum_{\wb{x} \in \wb{X}} h^d\Bigl[\bigl(u \ast \eta_h\bigr)(\wb{x})\Bigr]^2 \\
+	& \leq \sum_{\wb{x} \in \wb{X}}\int_{Q_h(\wb{x})} \bigl[u(y)\bigr]^2 \\
+	& = \norm{u}_{\Leb^2([0,1]^d)}^2.
+	\end{align*}
+	As a result,
+	\begin{align*}
+	\norm{\mathcal{K}_j(u_{\infty})}_{\Leb^2([0,1]^d)} & \leq \norm{\mathcal{K}_j(u_{\infty} - u_j)}_{\Leb^2([0,1]^d)} + \norm{\mathcal{K}_j(u_j)}_{\Leb^2([0,1]^d)} \\
+	& \leq \norm{u_\infty - u_j}_{\Leb^2([0,1]^d)} + 1/j \to 0
+	\end{align*}
+	as $j \to \infty$.
+On the other hand, $\mathcal{K}_j(u_{\infty}) \to u_{\infty}$ as $j \to \infty$.
+In light of the boundedness of $\eta$ and the boundedness of $u_{\infty}$ on $[0,1]^d$, we may apply the dominated convergence theorem and obtain that $\norm{\mathcal{K}_j(u_{\infty})}_{\Leb^2([0,1]^d)} \to \norm{u_{\infty}}_{\Leb^2([0,1]^d)}$.
+This establishes a contradiction, thus proving the claim.
+\end{proof}
+
+\subsection{Grid Sobolev seminorm}
+
+Actually, what I said before was wrong.
+Let me prove it.
+For notational convenience let $h = t^{-1}$.
+We write $\Delta = -\sum_{i = 1}^{d} \frac{\partial^2}{\partial e_i^2}$ for the continuum Laplace operator on $\Rd$.
+
+\begin{lemma} 
+	\label{lem:grid_sobolev_seminorm}
+	Let $f \in C_c^s(\Xset)$.
+	\begin{enumerate}
+		\item If $s$ is even, let $q = s/2$. We have
+		\begin{equation}
+		\label{eqn:grid_sobolev_seminorm_1}
+		\abs{\wb{L}^qf(\wb{x})} \leq c \norm{f}_{C^s(\Xset)} h^s
+		\end{equation}
+		for every $\wb{x} \in \wb{X}$.
+		\item If $s$ is odd, let $q = (s - 1)/2$. Then
+		\begin{enumerate}
+			\item If $\wb{x} \in \Xset_{q/t}$, then
+			\begin{equation}
+			\label{eqn:grid_sobolev_seminorm_2}
+			\abs{\wb{L}^qf(\wb{x}) - h^{s - 1} \Delta^q f(\wb{x})} \leq c \norm{f}_{C^s(\Xset)} h^s
+			\end{equation}
+			\item Otherwise if $\wb{x} \in \partial_{h}(\Xset)$, then
+			\begin{equation*}
+			\abs{\wb{L}^qf(\wb{x})} \leq c \norm{f}_{C^s(\Xset)} h^s
+			\end{equation*}
+		\end{enumerate} 
+	\end{enumerate}
+\end{lemma}
+\begin{proof}
+	We split our analysis into cases.
+
+	\paragraph{Proof of 1.}
+	Evaluated at an interior point $\wb{x} \in \Xset_r$, we have that
+	\begin{equation}
+	\label{eqn:grid_sobolev_seminorm_pf1}
+	\wb{L}f(\wb{x}) = \sum_{i = 1}^d \wb{D}_{e_i^2}f(\wb{x})
+	\end{equation}
+	where we use the notation $\wb{D}_{e_i^2}f(\wb{x}) := f(\wb{x} + he_i) + f(\wb{x} - he_i) - 2f(\wb{x})$.
+The second difference operator $\wb{D}_{e_i^2}$ zeros out the odd terms in a Taylor expansion as follows,
+	\begin{align*}
+	\wb{D}_{e_i^2}f(\wb{x}) & = \sum_{\alpha = 1}^{s - 1} \frac{\partial^{\alpha}}{\partial e_i^{\alpha}}f(\wb{x}) \Bigl[\bigl(he_i\bigr)^{\alpha} + \bigl(-he_i\bigr)^{\alpha}\Bigr] + R(\wb{x},i) \\
+	& = 2\sum_{\beta = 1}^{q - 1} h^{2\beta}\frac{\partial^{2\beta}}{\partial e_i^{2\beta}}f(\wb{x}) + R(\wb{x},i).
+	\end{align*}
+	Here $R(\wb{x},i)$ is the remainder term which satisfies
+	\begin{equation*}
+	\abs{R(\wb{x},i)} = \abs{\int_{-h}^{h}  \frac{\partial^{s}}{\partial e_i^{s}} f(\wb{x} + ye_i) y^{s - 1}\,dy} \leq 2 h^s \norm{f}_{C^s(\Xset)}
+	\end{equation*}
+	by the Holder property of $f$. Plugging back in to~\eqref{eqn:grid_sobolev_seminorm_pf1}, we obtain
+	\begin{equation*}
+	\wb{L}f(\wb{x}) = 2\sum_{i = 1}^{d} \biggl\{\sum_{\beta = 1}^{q - 1} h^{2\beta}\frac{\partial^{2\beta}}{\partial e_i^{2\beta}}f(\wb{x}) + R(\wb{x},i)\biggr\}
+	\end{equation*}
+	Repeatedly applying this estimate along with the triangle inequality gives that for any $\wb{x} \in \Xset_{qr}$,
+	\begin{equation}
+	\label{eqn:grid_sobolev_seminorm_pf2}
+	\abs{\wb{L}^qf(\wb{x}) - 2 \sum_{i \in [d]^q} \sum_{\beta \in [q - 1]^q} h^{2\abs{\beta}} \frac{\partial^{2\abs{\beta}}}{\partial e_{i_1}^{2\beta_1}\ldots \partial e_{i_q}^{2\beta_q} f(\wb{x})} } \leq c h^s\norm{f}_{C^s(\Xset)}
+	\end{equation}
+	where the interior sum over all multi-indices $\beta$ such that $\abs{\beta} \leq s$.
+Since $\abs{\beta}$ is at least $q$ for all $\beta \in [q]^q$, using the Holder property of $f$ for a second time we have
+	\begin{equation*}
+	h^{2\abs{\beta}} \abs{\frac{\partial^{2\abs{\beta}}}{\partial e_{i_1}^{2\beta_1}\ldots \partial e_{i_q}^{2\beta_q}} f(\wb{x})} \leq h^s \norm{f}_{C^s(\Xset)}
+	\end{equation*}
+	which along with~\eqref{eqn:grid_sobolev_seminorm_pf2} implies~\eqref{eqn:grid_sobolev_seminorm_1} for each $\wb{x} \in \Xset_{qh}$.
+
+	If instead $\wb{x} \in \partial_{qh}(\Xset)$, there exists some boundary point $x_0 \in \partial(\Xset)$ such that $\mathrm{dist}(\wb{x},x_0) \leq qh$.
+The order $\alpha$-terms, for $\alpha = 1,\ldots,s-1$, in the Taylor expansion of ${f}(\wb{x})$ about $f(x_0)$ are therefore zero, and
+	\begin{equation*}
+	\abs{f(\wb{x})} \leq (2qh)^s \norm{f}_{C^s(\Xset)}.
+	\end{equation*}
+	Applying similar logic to all $\wb{z} \in Q_{2qh}(\wb{z})$, we have that
+	\begin{equation*}
+	\abs{\wb{L}f(\wb{x})} \leq (2q)^d \cdot \max_{\wb{z} \in Q_{2qh}(\wb{z})} \Bigl\{\abs{f(\wb{z})}\Bigr\} \leq (2q)^{d + s} h^s \norm{f}_{C^s(\Xset)}.
+	\end{equation*}
+	Therefore~\eqref{eqn:grid_sobolev_seminorm_1} holds for all $\wb{x} \in \wb{X}$, regardless of their distance to the boundary.
+
+	\paragraph{Proof of 2a.}
+	We now deal with the case where $s$ is odd, and take $q = (s - 1)/2$. Similar reasoning to that of the preceding section implies that for any $\wb{x} \in \Xset_r$,
+	\begin{equation*}
+	\wb{L}f(\wb{x}) =  2\sum_{i = 1}^{d} \biggl\{\sum_{\beta = 1}^{q} h^{2\beta}\frac{\partial^{2\beta}}{\partial e_i^{2\beta}}f(\wb{x}) + R(\wb{x},i)\biggr\}
+	\end{equation*}
+	and additionally
+	\begin{equation}
+	\label{eqn:grid_sobolev_seminorm_pf3}
+	\abs{\wb{L}^qf(\wb{x}) - 2 \sum_{i \in [d]^q} \sum_{\beta \in [q]^q} h^{2\abs{\beta}} \frac{\partial^{2\abs{\beta}}}{\partial e_{i_1}^{2\beta_1}\ldots \partial e_{i_q}^{2\beta_q} f(\wb{x})} } \leq c h^s\norm{f}_{C^s(\Xset)}
+	\end{equation}
+	where as before $\abs{\beta} < s$ in the interior sum. Now, for the specific choice of $\beta_{\star} = (1,\ldots,1)$, we have
+	\begin{equation*}
+	\sum_{i \in [d]^q} h^{2\abs{\beta_{\star}}} \frac{\partial^{2q}}{\partial e_{i_1}^{2}\ldots \partial e_{i_q}^{2}}f(\wb{x}) = h^{{s - 1}} \Delta^q f(\wb{x}_i).
+	\end{equation*}
+	For all other $\beta \neq \beta_{\star} \in [q]^q$ satisfying $\abs{\beta} < s$ we have the same estimate as before,
+	\begin{equation*}
+	h^{2\abs{\beta}} \abs{\frac{\partial^{2\abs{\beta}}}{\partial e_{i_1}^{2\beta_1}\ldots \partial e_{i_q}^{2\beta_q}} f(\wb{x})} \leq h^s \norm{f}_{C^s(\Xset)}
+	\end{equation*}
+	and plugging back in to~\eqref{eqn:grid_sobolev_seminorm_pf3} establishes~\eqref{eqn:grid_sobolev_seminorm_2}.
+
+	\paragraph{Proof of 2b.}
+	The proof of 2b. follows along the same lines as the boundary case when $s$ was even.
+\end{proof}
+
+\section{\textcolor{red}{Old Stuff.}}
+To do so, we rely on Lemma~\ref{lem:grid_sobolev_seminorm}, which estimates on the absolute value of $L^q f$.
+When $s$ is even, the result follows immediately from~\eqref{eqn:grid_sobolev_norm_higher_order} in that Lemma,
+\begin{equation*}
+f^T \wb{L}^s f = \sum_{\wb{x} \in \wb{X}} \bigl(\wb{L}^q f(\wb{x})\bigr)^2 \leq c n \norm{f}_{C^s(\Xset)}^2 t^{-2s}.
+\end{equation*}
+When $s$ is even, the result follows by distinguishing between points close to the boundary and points sufficiently on the interior,
+\begin{align*}
+f^T \wb{L}^s f  & = \sum_{\wb{x}_i,\wb{x}_j \in \wb{X}} \bigl(\wb{L}^q f(\wb{x}_i) - \wb{L}^q f(\wb{x}_i)\bigr)^2 \1\{\norm{\wb{x}_i - \wb{x}_j}_1 \leq h\} \\
+& \leq c n t^{-2s} \norm{f}_{C^s(\Xset)}^2 + \sum_{\wb{x}_i,\wb{x}_j \in \wb{X} \cap \Xset_{qh}} \bigl(\wb{L}^q f(\wb{x}_i) - \wb{L}^q f(\wb{x}_j)\bigr)^2 \1\{\norm{\wb{x}_i - \wb{x}_j}_1 \leq h\} \\
+& \leq c n t^{-2s} \norm{f}_{C^s(\Xset)}^2 + h^{2(s - 1)}\sum_{\wb{x}_i,\wb{x}_j \in \wb{X} \cap \Xset_{qh}} \bigl(\Delta^q f(\wb{x}_i) - \Delta^q f(\wb{x}_j)\bigr)^2 \1\{\norm{\wb{x}_i - \wb{x}_j}_1 \leq h\} \\
+& \leq c n t^{-2s} \norm{f}_{C^s(\Xset)}^2
+\end{align*}
+where the last inequality follows from applying the Holder condition to $\Delta^q f$.
+
 \end{document}

### 2020-08-19 notes/weekly_notes/3.26.20.NotesForWeek.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\makeatletter
\newcommand{\vast}{\bBigg@{4}}
\newcommand{\Vast}{\bBigg@{5}}
\makeatother

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{\mathcal{L}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LS}{\mathrm{LS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 3/19/20 - 3/26/20}
\author{Alden Green}
\date{\today}
\maketitle

Let $G_0 = ([\mc{M}],E_0)$ be a graph on $\mc{M} \geq 1$ vertices. Let $N_1, \ldots, N_{\mc{M}}$ each be positive integers. The \textcolor{red}{(Alden product)} graph $H_0$ on $n = \sum N_m$ nodes is defined as
\begin{equation*}
H_0 = \biggl(\bigcup_{m = 1}^{\mc{M}} \bigcup_{i = 1}^{N_m} (m,i),~~ F_0\biggr),~~\textrm{where}~ (\ell,i) \sim (m,j)~~\textrm{in $H_0$ if $\ell \sim m$ in $G_0$.}  
\end{equation*}
Let $\lambda_k(H_0)$ denote the $k$th smallest eigenvalue of the Laplacian $L_{H_0}$, and let
\begin{equation*}
N_{\min} = \min_{m \in [\mc{M}]} N_m,~~N_{\max} = \min_{m \in [\mc{M}]} N_m
\end{equation*} We wish to prove the following Lemma
\begin{lemma}
	\label{lem:alden_product_eigenvalues}
	For $k \in [n]$,
	\begin{equation}
	\label{eqn:alden_product_eigenvalue}
	\lambda_k(H_0) \geq \frac{\deg_{\min}(G_0)N_{\min}^3}{\deg_{\max}(G_0)N_{\max}^2}  \cdot
	\begin{cases}
	\lambda_k(G_0),& ~~\textrm{if $k \in [\mc{M}]$} \\
	1,& ~~\textrm{otherwise.}
	\end{cases}
	\end{equation}
\end{lemma}

Our proof will involve a series of matrix and graph comparisons. For a graph $G$ on $M$ vertices, let $N_G$ be the normalized Laplacian matrix, with associated eigenvalues $\lambda_1(N_G) \leq \cdots \leq \lambda_M(N_G)$. Let  $\sigma_1,\ldots,\sigma_{\mc{M}}$ be contractions from $[N_{\max}]$ to $[N_{1}],\ldots,[N_{\mc{M}}]$, respectively, to be defined later. The mappings $\sigma_m$ and the graph $H_0$ jointly induce a weighted graph 
\begin{equation}
\label{eqn:alden_product_eigenvalue_0}
\wb{H}_0 = \biggl([\mc{M}] \times [N_{\max}], W_0\biggr),~~ \textrm{where $W_0[(m,j),(m',j')] = \frac{ \1\bigl\{(m,\sigma_m(j)) \sim (m',\sigma_{m'}(j'))~\textrm{in $H$} \bigr\}}{|\{\ell:\sigma_m(\ell) = \sigma_m(j)\}| \cdot |\{\ell':\sigma_{m'}(\ell') = j' \}|}$}
\end{equation}

\textcolor{red}{(TODO)}: Define $\wb{G}_0$ here.

We shall proceed according to the following steps:
\begin{enumerate}
	\item For each $k \in [n]$, 
	\begin{equation}
	\label{eqn:alden_product_eigenvalue_pf1}
	\lambda_k(H_0) \geq \deg_{\min}(H_0) \cdot \lambda_k(N_{H_0})
	\end{equation}
	\item For each $k \in \bigl[n\bigr]$,
	\begin{equation}
	\label{eqn:alden_product_eigenvalue_pf2}
	\lambda_k(N_{H_0}) \geq \lambda_k(N_{\wb{H}_0})
	\end{equation}
	\item For each $k \in [\mc{M} N_{\max}]$, 
	\begin{equation}
	\label{eqn:alden_product_eigenvalue_pf3}
	\lambda_k(N_{\wb{H}_0}) \geq \frac{N_{\min}^2}{N_{\max}^2} \cdot \lambda_k(N_{\wb{G}_0})
	\end{equation}
	\item For each $k \in [\mc{M}]$,
	\begin{equation}
	\label{eqn:alden_product_eigenvalue_pf4}
	\wt{\lambda}_k(\wb{G}_0) = \wt{\lambda}_k(G_0)
	\end{equation}
	Otherwise if $k > \mc{M}$, $\wt{\lambda}_k(\wb{G}_0) = 1$.
	\item For each $k \in [\mc{M}]$,
	\begin{equation*}
	\wt{\lambda}_k(G_0) \geq \frac{\lambda_k(G_0)}{\deg_{\max}(G)}.
	\end{equation*}
\end{enumerate}

\textcolor{red}{(TODO): Fill in proof.}

\paragraph{Step 1: Moving to Normalized Laplacian.}
For simplicity, in this section we will deal with an arbitrary $G = ([n], E)$, and show
\begin{equation*}
\lambda_k(G) \geq \deg_{\min}(G) \cdot \lambda_k(N_G)
\end{equation*}
for all $k \in [n]$. If $\deg_{\min}(G) = 0$, the statement is trivially obvious, and so we will suppose without loss of generality that $\deg_{\min}(G) > 0$.

Let $L = D - A$, where $A = A(G)$ is the adjacency matrix of $G$, $D := D(G)$ is the degree matrix associated with $G$, and $L$ is therefore the Laplacian of $G$. Let $N = D^{-1/2} L D^{-1/2}$ be the normalized Laplacian of $G$. Using the Courant-Fischer min max theorem, and letting $D^{-1/2} V = \{D^{-1/2} v: v\in V\}$ for any subspace $V \subset \Reals^n$, we have
\begin{align*}
\lambda_k(G) & = \min_{V} \Bigl\{ \max_{v \in V} \frac{v^T L v}{v^T v} \Bigr\} \\
& = \min_{V} \Bigl\{ \max_{u \in D^{-1/2} V} \frac{u^T N u}{u^T D^{-1} u} \Bigr\} \\
& \geq \deg_{\min}(G) \min_{V} \Bigl\{ \max_{u \in D^{-1/2} V} \frac{u^T N u}{u^T u} \Bigr\}
\end{align*}
where the minimum is always over all $k$ dimensional subspaces of $\Reals^n$ and the second line follows upon substituting $u = D^{1/2} v$. Since every vertex has non-zero degree, both $D^{1/2}$ and $D^{-1/2}$ are full rank matrices, and $\dim(D^{-1/2} V) = \dim(V)$ for all subspaces $V$. Hence, we have
\begin{equation*}
\min_{V} \Bigl\{ \max_{u \in D^{-1/2} V} \frac{u^T N u}{u^T u} \Bigr\} = \min_{U} \Bigl\{ \max_{u \in U} \frac{u^T N u}{u^T u} \Bigr\} = \lambda_k(N_G).
\end{equation*} 
Choosing $G = H_0$, we obtain~\eqref{eqn:alden_product_eigenvalue_pf1}.

\paragraph{Step 2: Moving to a weighted graph.}

By definition, the graph $H_0$ is a contraction of $\wb{H}_0$. Moreover, for any vertices $(m,j)$ and $(m',j')$ which are contracted together, $m = m'$ and $\sigma_m(j) = \sigma_m(j')$, so the two vertices have all the same edge weights in $\wb{H}_0$. The eigenvalue inequality~\eqref{eqn:alden_product_eigenvalue_pf2} then follows from Lemma~\ref{lem:contraction}.

\paragraph{Step 3: Moving to unweighted tensor product graph.}
We first lower bound $N_{\wb{H}_0}$ by an unweighted tensor product graph $N_{\wb{G}_0}$, where $\wb{G}_0$ is defined as
\begin{equation*}
\wb{G}_0 = \bigl([\mc{M}] \times [N_{\max}], E(\wb{G}_0)\bigr),~~\textrm{where}~~ (m,j) \sim (m',j')~ \textrm{in $\wb{G}_0$ if}~ m \sim m'~\textrm{in $G_0$.} 
\end{equation*}
The graph $\wb{H}_0$ has the same adjacency structure as $\wb{G}_0$, but edges with weights between $0$ and $1$. To have our lower bound be sufficiently large, we would like to make the minimum of these weights large. We achieve by this ensuring the maps $\sigma_m$ do not map too many vertices in $[\mc{M}] \times [N_{\max}]$ to the same vertex in $V(H_0)$. Formally, we let
\begin{equation*}
\sigma_m\bigl(i\bigr) := i \mod N_{m}
\end{equation*}
Then, for any $(m,i) \in [\mc{M}] \times [N_{\max}]$, clearly $\abs{\{\ell:\sigma_m(\ell) = \sigma_m(j)\}} \leq N_{\max} / N_{\min}$. Recalling the definition of the weight matrix $W_0$ in~\eqref{eqn:alden_product_eigenvalue_0}, by Lemma~\ref{lem:weighted_graph_comparison}, we have that
\begin{equation*}
\lambda_k(N_{\wb{H}_0}) \geq \frac{N_{\min}^2}{N_{\max}^2} \lambda_k(N_{\wb{G}_0}).
\end{equation*}

\paragraph{Step 4: Completing the proof.}
We have that $\wb{G}_0 = G_0 \otimes K_{N_{\max}}$, and we may therefore characterize the spectrum $\wt{\Lambda}(\wb{G}_0)$ by $\wt{\Lambda}(G_0)$ and $\wt{\Lambda}(K_{N_{\max}})$ using Lemma~\ref{lem:tensor}. The latter spectrum is simply
\begin{equation*}
\lambda_1(N_{K_{N_{\max}}}) = 0, \lambda_2(N_{K_{N_{\max}}}),\ldots,\lambda_N(N_{K_{N_{\max}}}) = 1,
\end{equation*}
and therefore by Lemma~\ref{lem:tensor}
\begin{equation}
\wt{\lambda}_k(\wb{G}_0) =
\begin{cases*}
\wt{\lambda}_k(G_0),& ~~\textrm{for $k = 1,\ldots,\mc{M}$} \\
1,& ~~ \textrm{for $k = \mc{M} + 1,\ldots, \mc{M} \cdot N_{\max}$.}
\end{cases*}
\end{equation}

\section{Additional Theory}

\subsection{Contractions}

Let $H = ([n + m], W_H)$ be an arbitrary weighted graph on $n + m$ vertices. Any mapping $\sigma: [n + m] \to [n]$ induces a graph $G = ([n], W_G)$ with weights
\begin{equation*}
W_G[k,k'] = \sum_{\ell: \sigma(\ell) = k} \sum_{\ell': \sigma(\ell') = k'} W_H[\ell,\ell']
\end{equation*}
which we call the contraction of $H$ induced by $\sigma$.

\begin{lemma}
	\label{lem:contraction}
	Suppose that the graph $H$ and contraction $\sigma$ satisfy the following property: for all $\ell, \ell'$ such that $\sigma(\ell) = \sigma(\ell')$, $W_H[\ell,\cdot] = W_H[\ell',\cdot]$. Then, 
	\begin{equation*}
	\lambda_k(N_G) \geq \lambda_k(N_{H}),~~\textrm{for all $k \in [n]$.}
	\end{equation*}
\end{lemma}
\begin{proof}
	The following two facts are key consequences of the assumption that $W_H[\ell,\cdot] = W_H[\ell',\cdot]$ for all $\sigma(\ell) = \sigma(\ell')$. First, for any $i$ and $i' \in [n + m]$,
	\begin{align*}
	W_G[\sigma(i),\sigma(i')] & = \sum_{\ell: \sigma(\ell) = \sigma(i)} \sum_{\ell': \sigma(\ell') = \sigma(i')} W_H[\ell,\ell'] \\
	& = \sum_{\ell: \sigma(\ell) = \sigma(i)} \sum_{\ell': \sigma(\ell') = \sigma(i')} W_H[i,i'] = W_H[i,i'] \cdot N_{\sigma}(i) N_{\sigma}(i').
	\end{align*}
	Second, for any $\ell \in [n + m]$,
	\begin{align*}
	\deg_H(\ell) & = \sum_{i = 1}^{n + m} W_H[i,\ell] \\
	& = \sum_{i = 1}^{n + m} \frac{W_G\bigl[\sigma(i),\sigma(\ell)\bigr]}{N_{\sigma}(i) N_{\sigma(\ell)}} \\
	& = \sum_{j = 1}^{n} \frac{W_G\bigl[j,\ell\bigr]}{N_{\sigma}(\ell)} \\
	& = \frac{\deg_G(\sigma(\ell))}{N_{\sigma}(\ell)}.
	\end{align*}

	Now, let $v$ be an eigenvector of $N_G$, meaning there exists $\lambda > 0$ such that
	\begin{equation*}
	N_G v = \lambda v
	\end{equation*}
	Define $u: [n + m] \to \Reals$ to be the vector
	\begin{equation*}
	u(\ell) = \frac{v(\sigma(\ell))}{\sqrt{N_{\sigma}(\ell)}},~~\textrm{where $N_{\sigma}(\ell) = \Bigl|\{\ell':\sigma(\ell') = \sigma(\ell)\}\Bigr|$}
	\end{equation*}
	The following manipulations show that $u$ is an eigenvector of $N_H$, with eigenvalue $\lambda$. 
	\begin{align*}
	\Bigl(N_H u\Bigr)(\ell) & = \sum_{i = 1}^{n + m} \Bigl\{ \frac{u(\ell)}{\deg_H(\ell)} - \frac{u(i)}{\sqrt{\deg_H(\ell) \deg_H(i)}} \Bigr\} W_{H}[\ell,i] \\
	 & = \sum_{i = 1}^{n + m} \Bigl\{ \frac{v\bigl(\sigma(\ell)\bigr)}{\sqrt{N_{\sigma}(\ell)}\deg_H(\ell)} - \frac{v\bigl(\sigma(i)\bigr)}{\sqrt{N_{\sigma}(i)\deg_H(\ell) \deg_H(i)}} \Bigr\} W_{H}[\ell,i] \\
	& = \sum_{i = 1}^{n + m} \Bigl\{ \frac{\sqrt{N_{\sigma}(\ell)} v\bigl(\sigma(\ell)\bigr)}{\deg_G\bigl(\sigma(\ell)\bigr)} - \frac{\sqrt{N_{\sigma}(\ell)} v\bigl(\sigma(i)\bigr)}{\sqrt{\deg_G\bigl(\sigma(\ell)\bigr) \deg_G\bigl(\sigma(i)\bigr)}} \Bigr\} W_{H}[\ell,i]  \\
	& = \sum_{i = 1}^{n + m} \Bigl\{ \frac{\sqrt{N_{\sigma}(\ell)} v\bigl(\sigma(\ell)\bigr)}{\deg_G\bigl(\sigma(\ell)\bigr)} - \frac{\sqrt{N_{\sigma}(\ell)} v\bigl(\sigma(i)\bigr)}{\sqrt{\deg_G\bigl(\sigma(\ell)\bigr) \deg_G\bigl(\sigma(i)\bigr)}} \Bigr\} \frac{W_{G}\Bigl[\sigma(\ell),\sigma(i)\Bigr]}{N_{\sigma}(\ell) N_{\sigma}(i)} \\
	& = \frac{1}{\sqrt{N_{\sigma}(\ell)}}\sum_{i = 1}^{n + m} \Bigl\{ \frac{ v\bigl(\sigma(\ell)\bigr)}{\deg_G\bigl(\sigma(\ell)\bigr)} - \frac{ v\bigl(\sigma(i)\bigr)}{\sqrt{\deg_G\bigl(\sigma(\ell)\bigr) \deg_G\bigl(\sigma(i)\bigr)}} \Bigr\} \frac{W_{G}\Bigl[\sigma(\ell),\sigma(i)\Bigr]}{N_{\sigma}(i)} \\
	& \overset{(i)}{=} \frac{1}{\sqrt{N_{\sigma}(\ell)}}\sum_{j = 1}^{n} \Bigl\{ \frac{ v\bigl(\sigma(\ell)\bigr)}{\deg_G\bigl(\sigma(\ell)\bigr)} - \frac{ v\bigl(j\bigr)}{\sqrt{\deg_G\bigl(\sigma(\ell)\bigr) \deg_G\bigl(j\bigr)}} \Bigr\} W_{G}\Bigl[\sigma(\ell),j\Bigr] \\
	& = \frac{1}{\sqrt{N_{\sigma}(\ell)}} \lambda v\bigl(\sigma(\ell)\bigr) = \lambda u(\ell),
	\end{align*}
	where $(i)$ follows from the substitution $j = \sigma(i)$.

	Therefore every eigenvalue of $G$ is also an eigenvalue of $H$. It follows immediately that the $k$th smallest eigenvalue of $H$ must be no greater than the $k$th smallest eigenvalue of $G$.
\end{proof}

\subsection{Weighted graphs}

\begin{lemma}
	\label{lem:weighted_graph_comparison}
	Let $G = ([n],E)$ be an unweighted and connected graph, and let $A$ be an $n \times n$ symmetric matrix with entries $0 < A_{ij} < 1$. Let $H = ([n],W)$ be a weighted graph with weights
	\begin{equation*}
	W_{ij} = A_{ij} \times \1\{(i,j) \in G\}.
	\end{equation*}
	Then, 
	\begin{equation*}
	\lambda_k(N_H) \geq \min \{A_{ij}\}  \cdot \lambda_k(N_G)
	\end{equation*}
	for all $k = 1,\ldots,n$.
\end{lemma}
\begin{proof}
	Note that since $A$ has strictly positive entries and $G$ is connected, the degree of every vertex $i \in [n]$ is positive in both $G$ and $H$; thus $D_H$ and $D_G$ are full rank, and so is $D_G^{1/2}D_H^{-1/2}$. By the Courant-Fischer Theorem,
	\begin{align*}
	\lambda_k(N_H) & = \min_V \biggl\{ \max_{v} \Bigl\{\frac{v^T N_H v}{v^T v}: v \in V~\textrm{and}~ v \neq 0\Bigr\} : \dim(V) = k\biggr\} \\
	& = \min_V \biggl\{ \max_{u} \Bigl\{\frac{u^T L_H u}{u^T D_H u}: u \in D_H^{-1/2} V~\textrm{and}~ u \neq 0\Bigr\} : \dim(V) = k\biggr\} \tag{substituting $u = D_H^{-1/2}v$}\\ 
	& \geq \min \{A_{ij} \} \cdot \min_{V} \biggl\{ \max_{u} \Bigl\{\frac{u^T L_G u}{u^T D_G u}: u \in D_H^{-1/2} V~\textrm{and}~ u \neq 0\Bigr\} : \dim(V) = k\biggr\} \\
	& = \min \{A_{ij} \} \cdot \min_{V} \biggl\{ \max_{w} \Bigl\{\frac{w^T N_G w}{w^Tw}: w \in D_G^{1/2} D_H^{-1/2} V~\textrm{and}~ u \neq 0\Bigr\} : \dim(V) = k\biggr\} \tag{substituing $w = D_G^{1/2} u$} \\
	& = \min \{A_{ij} \} \cdot \lambda_k(N_G),
	\end{align*}
	where the last inequality follows from Lemma~\ref{lem:variational}.
\end{proof}

\subsection{Tensor product graphs}
For an unweighted graph $G = ([n],E)$, the random walk matrix $P_G$ has entries
\begin{equation*}
(P_G)_{ij} = \frac{1}{\deg_G(i)} \1\{(i,j) \in E\}.
\end{equation*}
Recall that the spectrum $\Lambda(P_G) = 1 - \Lambda(N_G)$ (see, e.g. \textcolor{red}{(von Luxburg)}).

For graphs $G_1 = ([N],E_1)$ and $G_2 = ([M],E_2)$, the tensor $H = G_1 \otimes G_2$ is defined over the vertex set $[N] \times [M]$, with an edge between $(i,k)$ and $(j,l)$ if and only if $i$ is connected to $j$ in $G_1$ and $k$ is connected to $l$ in $G_2$.
\begin{lemma}
	\label{lem:tensor}
	Let $G_1 = ([N],E_1)$ and $G_2 = ([M],E_2)$ be unweighted graphs, and let $H = G_1 \otimes G_2$. Then, the spectrum
	\begin{equation*}
	\Lambda(N_H) = \Bigl\{\lambda_k(N_{G_1}) + \lambda_k(N_{G_2}) - \lambda_k(N_{G_1}) \cdot \lambda_{\ell}(N_{G_1}): (k,\ell) \in [N] \times [M] \Bigr\}
	\end{equation*}
\end{lemma}
\begin{proof}
	We will abbreviate $P_1 := P_{G_1}$ and $P_2 := P_{G_2}$. Let $v_k$ and $u_{\ell}$ satisfy
	\begin{equation*}
	P_1 v_{k} = \lambda_k v_k,~~ P_2 u_{\ell} = \lambda_{\ell} u_{\ell}.
	\end{equation*}

	Let $w: \Reals^{N \times M}$ be defined by $w_{ij}  = v_{k,i} u_{\ell,j}$. We will show that $w$ is an eigenvector of $P_H$ satisfying 
	\begin{equation*}
	P_H w = \lambda_k \lambda_{\ell} \cdot w.
	\end{equation*}
	To see this, note that
	\begin{align*}
	\deg_H\bigl((i,j)\bigr) & = \sum_{i' = 1}^{N} \sum_{j' = 1}^{M} \1\Bigl\{(i,j) \sim (i',j') ~\textrm{in}~ H \Bigr\} \\
	& = \sum_{i' = 1}^{N} \sum_{j' = 1}^{M} \1\Bigl\{i \sim i' ~\textrm{in}~ G_1 \Bigr\} \1\Bigl\{j \sim j' ~\textrm{in}~ G_2 \Bigr\} \\ 
	& = \deg_{G_1}(i) \deg_{G_2}(j),
	\end{align*}
	and therefore
	\begin{align*}
	\bigl(P_Hw\Bigr)_{ij} & = \sum_{i' = 1}^{N} \sum_{j' = 1}^{M} \frac{w_{i'j'}}{\deg_H(i',j')} \1\Bigl\{(i,j) \sim (i',j') ~\textrm{in}~ H \Bigr\} \\
	& = \sum_{i' = 1}^{N} \sum_{j' = 1}^{M} \frac{v_{ki'} u_{\ell j'}}{\deg_{G_1}(i') \deg_{G_2}(j')} \1\Bigl\{i \sim i' ~\textrm{in}~ G_1 \Bigr\} \1\Bigl\{j \sim j' ~\textrm{in}~ G_2 \Bigr\} \\
	& = \biggl(\sum_{i' = 1}^{N} \sum_{j' = 1}^{M} \frac{v_{ki'}}{\deg_{G_1}(i')} \1\Bigl\{i \sim i' ~\textrm{in}~ G_1 \Bigr\} \biggr) \biggl(\sum_{i' = 1}^{N} \sum_{j' = 1}^{M} \frac{u_{kj'}}{\deg_{G_2}(j')} \1\Bigl\{j \sim j' ~\textrm{in}~ G_1 \Bigr\} \biggr) \\
	& = \lambda_k \lambda_{\ell} v_{ki} u_{\ell j}.
	\end{align*}

	This characterizes the spectrum $\Lambda(P_H)$. The claim of Lemma~\ref{lem:tensor} follows upon recalling that the spectrum $\Lambda(N_G) = 1 - \Lambda(N_G)$ for $G = H$, $G = G_1$ and $G = G_2$, so that
	\begin{align*}
	\lambda_{k,\ell}(N_H) & = 1 - \lambda_{k,\ell}(P_H) \\
	& = 1 - \lambda_{k}(P_{G_1}) \lambda_{\ell}(P_{G_2}) \\
	& = 1 - \bigl(1 - \lambda_{k}(N_{G_1})\bigr) \bigl(1 - \lambda_{\ell}(N_{G_2})\bigr) \\
	& = \lambda_{k}(N_{G_1}) + \lambda_{\ell}(N_{G_2}) - \lambda_{k}(N_{G_1}) \lambda_{\ell}(N_{G_2}).
	\end{align*}
\end{proof}

\subsection{Variational lemmas}

We will use the following fact repeatedly. We state and prove it formally as a sanity check. For a symmetric $n \times n$ matrix $A$, and a non-zero vector $v \in \Reals^n$, the Rayleigh quotient is 
\begin{equation*}
R_A(v) = \frac{v^T A v}{v^T v}
\end{equation*}
We let $\lambda_1(A) \leq \ldots \leq \lambda_n(A)$ be the eigenvalues of $A$, sorted in ascending order. For a subspace $V \subseteq \Reals^n$ and operator $D:\Reals^n \to \Reals^n$, let $DV := \set{Dv: v \in V}$.
\begin{lemma}
	\label{lem:variational}
	Let $A$ be an $n \times n$ matrix, and let $D: \Reals^n \to \Reals^n$ be a full rank linear operator. Then,
	\begin{equation*}
	\lambda_k(A) = \min_{V}\Bigl\{ \max_{v}\bigl\{R_{A}(v): v \in DV~\textrm{and}~v \neq 0\bigr\}: \dim(V) = k \Bigr\}
	\end{equation*}
\end{lemma}
\begin{proof}
	We know from the Courant-Fischer Theorem that 
	\begin{equation*}
	\lambda_k(A) = \min_{V}\Bigl\{ \max_{v}\bigl\{R_{A}(v): v \in V~\textrm{and}~v \neq 0\bigr\}: \dim(V) = k \Bigr\}
	\end{equation*}
	Let $(v_k,V_k)$ satisfy
	\begin{equation*}
	R_A(v_k) = \lambda_k(A), ~~\textrm{and}~~ v_k = \argmax_{v} \Bigl\{ R_A(v): v \in V_k~\textrm{and}~v \neq 0 \Bigr\}
	\end{equation*}
	Now, let $U_{\ast} = D^{-1}V_k$; since $D$ is a full rank operator $U_{\ast}$ is well-defined and $\dim(U_{\ast}) = \dim(V_k) = k$. Clearly $V_k = DU_{\ast}$, and therefore $v_k \in DU_{\ast}$. As a result
	\begin{align*}
	\min_{V}\Bigl\{ \max_{v}\bigl\{R_{A}(v): v \in DV~\textrm{and}~v \neq 0\bigr\}: \dim(V) = k \Bigr\} & \leq \max_{v}\bigl\{R_{A}(v): v \in DU_{\ast}~\textrm{and}~v \neq 0\bigr\} \\
	& = \max_{v}\bigl\{R_{A}(v): v \in V_k~\textrm{and}~v \neq 0\bigr\} \\
	& = R_A(v_k) \\
	& = \lambda_k(A).
	\end{align*}

	On the other hand, let
	\begin{equation*}
	U_k := \argmin_U \Bigl\{\max_{v} \bigl\{R_A(v): v \in DU~\textrm{and}~ v \neq 0\bigr\}: \dim(U) = k\Bigr\}
	\end{equation*}
	and let $V_{\ast} = DU_{k}$. Since $D$ is full rank $\dim(V_{\ast}) = \dim(U_{\ast}) = k$, and therefore
	\begin{align*}
	\lambda_k(A) & = \min_{V} \Bigl\{\max\bigl\{R_A(v): v \in V~\textrm{and}~ v \neq 0 \bigr\}: \dim(V) = k\Bigr\} \\
	& \leq \max\bigl\{R_A(v): v \in V_{\ast}~\textrm{and}~v \neq 0 \bigr\} \\
	& = \max\bigl\{R_A(v): v \in DU_{k}~\textrm{and}~v \neq 0 \bigr\} \\
	& = \min_U \Bigl\{\max_{v} \bigl\{R_A(v): v \in DU~\textrm{and}~v \neq 0\bigr\}: \dim(U) = k\Bigr\}
	\end{align*}
\end{proof}

Among other things, Lemma~\ref{lem:variational} allows us to compare the spectrum of $N_G$ and $L_G$, in terms of the minimum and maximum degree of $G$.
\begin{lemma}
	Let $G = ([N],E)$ be an unweighted connected graph. Then,
	\begin{equation*}
	\frac{{\lambda}_k(G)}{\deg_{\max}(G)} \leq \wt{\lambda}_k(G)  \leq \frac{{\lambda}_k(G)}{\deg_{\min}(G)}
	\end{equation*}
\end{lemma}
\begin{proof}
	The following manipulations establish the lower bound,
	\begin{align*}
	\lambda_k(N_G) & = \min_V \biggl\{ \max_{v} \Bigl\{\frac{v^T N_G v}{v^T v}: v \in V~\textrm{and}~ v \neq 0 \Bigr\}: \dim(V) = k \biggr\} \\
	& = \min_V \biggl\{ \max_{u} \Bigl\{\frac{u^T L_G u}{u^T D_G u}: u \in D^{-1/2} V~\textrm{and}~ u \neq 0 \Bigr\}: \dim(V) = k \biggr\} \\
	& \geq \frac{1}{\deg_{\max}(G)} \cdot \min_V \biggl\{ \max_{u} \Bigl\{\frac{u^T L_G u}{u^T u}: u \in D^{-1/2} V~\textrm{and}~ u \neq 0 \Bigr\}: \dim(V) = k \biggr\} \\
	& = \frac{\lambda_k(G)}{\deg_{\max}(G)},
	\end{align*}
	where the last inequality follows by Lemma~\ref{lem:variational}. The upper bound follows by similar steps, upon replacing $\deg_{\max}(G)$ by $\deg_{\min}(G)$ in the previous expression and reversing the inequality.
\end{proof}

\end{document}

### 2020-08-19 notes/weekly_notes/4.9.20.NotesForWeek.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\makeatletter
\newcommand{\vast}{\bBigg@{4}}
\newcommand{\Vast}{\bBigg@{5}}
\makeatother

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}
\newcommand{\Lap}{{\bf L}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\LE}{\mathrm{LE}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 4/3/20 - 4/9/20}
\author{Alden Green}
\date{\today}
\maketitle

Suppose we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to the following random design regression model. The design points $X_1,\ldots,X_n \in \Xset$ are independently sampled from a distribution $P$, and the responses
\begin{equation*} 
Y_i = f_0(X_i) + \varepsilon_i
\end{equation*}
where $f_0: \Xset \to \Reals$ is the unknown regression function and $\varepsilon_i$ are independent $N(0,1)$ noise samples. Our goal is to test
\begin{equation*}
\mathbf{H}_0: f_0 = 0,~~ \mathbf{H}_a: f_0 \neq 0.
\end{equation*}
Our test statistic $T = \norm{\wh{f}}_n^2$ will be a plug-in estimator of the $\Leb_2$-norm of $f_0$. The estimator $\wh{f}$ of the function $f_0$ will be a truncated-series estimator, using eigenvectors of a graph Laplacian. We now define the graph $\wb{G}_{n,r}$ we will build over the design points ${\bf X} = X_1,\ldots,X_n$, which we term the \emph{histogram lattice}.

\paragraph{Histogram Lattice.}

Let $0 < r < 1$, and let $M = 1/r$. Let
\begin{equation*}
Z = \Bigl\{ \frac{1}{M}\bigl(2m_1 - 1,\ldots,2m_d - 1\bigr): m \in [M]^d \Bigr\}
\end{equation*}
be a set of evenly spaced grid points. (We will always assume $M$ is an integer, merely to simplify some notational and definitional details.) A natural graph associated with $Z$ is the lattice
\begin{equation*}
\wb{G} = (Z,\wb{E}),~~ (z,z') \in \wb{E} ~\textrm{if}~ \norm{z - z'}_2 = r
\end{equation*}
Let $\Pi: [0,1]^d \to Z$ map points $x \in [0,1]^d$ to the nearest grid points $z \in Z$,
\begin{equation*}
\Pi(x) := \argmin_{z \in Z}~ \norm{z - x}_2 
\end{equation*}
The \emph{histogram lattice} $\wb{G}_{n,r} = ({\bf X}, \wb{E}_{n,r})$ is a product graph induced by $\wb{G}$ and the mapping $\Pi$, with edges 
\begin{equation*}
(X_i,X_j) \in \wb{E}_{n,r} ~\textrm{if}~ \Bigl(\Pi(X_i),\Pi(X_j)\Bigr) \in \wb{E}.
\end{equation*}

The matrix $\wb{L}_{n,r}$ is the graph Laplacian of $\wb{G}_{n,r}$ and $(\lambda_1,v_1),\ldots,(\lambda_n,v_n)$ are the eigenvector/eigenvalue pairs of $\wb{L}_{n,r}$, defined by the equation
\begin{equation*}
\wb{L}_{n,r} v_k = \lambda_k v_k,~~ \norm{v_k}_n^2 = 1.
\end{equation*}
For a positive integer $\kappa$, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ is defined as 
\begin{equation*}
\wh{f}_{\LE} := \sum_{k = 1}^{\kappa} \dotp{Y}{v_k}_n v_k.
\end{equation*}
and the resulting test statistic is thus
\begin{equation*}
\wh{T}_{\LE} = \sum_{k = 1}^{\kappa} \bigl(\dotp{Y}{v_k}_n\bigr)^2.
\end{equation*}

\section{Approximation Error}

We make some assumptions on the function $f_0$ and the distribution $P$.

\begin{enumerate}[label=(A\arabic*)]
	\item 
	\label{asmp:smooth_function}
	$f_0 \in C^{s}(\Xset;B)$ for some $s > 0$. If $s > 1$, then $f$ is also compactly supported on a strict subset of $\Xset$.

	\item 
	\label{asmp:smooth_density}
	$P$ admits a density $p$ with respect to the Lebesgue measure on $\Reals^d$. The density $p \in C^{}(\Xset;p_{\max})$, for some $k > 0$.
\end{enumerate}

\begin{theorem}
	\label{thm:approximation_error}
	Suppose assumptions~\ref{asmp:smooth_function} and~\ref{asmp:smooth_density} are satisfied for some $s \geq 1$, and $k = s - 1$. Then, there exists a constant $c$ such that 
	\begin{equation*}
	\sum_{k = 1}^{\kappa} \bigl(\dotp{v_k}{f_0}_n\bigr)^2 \geq c \cdot \norm{f_0}_2^2 - \kappa^{-2/d}
	\end{equation*}
\end{theorem}

\section{Proof of Theorem~\ref{thm:approximation_error}}

Let $\wb{f}_0: {\bf X} \to \Reals$ be the histogram estimate of $f_0$, defined as
\begin{equation*}
\wb{f}_0(X) = \frac{1}{\abs{Q(\Pi(X))}} \sum_{i = 1}^{n} f_0(X_i) \cdot \1\Bigl\{\pi(X_i) = \pi(X)\Bigr\}
\end{equation*}
We shall proceed according to the following steps.
\begin{enumerate}
	\item Under the assumption~\ref{asmp:smooth_density}, for any $(\log(n)/n)^{1/d} \leq r$, there exist constants $c$ and $C$ such that
	\begin{equation}
	\label{eqn:approximation_error_pf1}
	c \cdot \min\Bigl\{nr^{d+2}\kappa^{2/d}, \deg_{\min}(\wb{G}_{n,r})\Bigr\} \leq \lambda_{\kappa}(\wb{G}_{n,r}) \leq C \cdot nr^{d + 2}\kappa^{2/d}
	\end{equation}
	and
	\begin{equation}
	\label{eqn:approximation_error_pf2}
	c\cdot nr^{d} \leq \deg_{\min}(\wb{G}_{n,r}) \leq C \cdot nr^d
	\end{equation}
	with probability at least $1 - \textcolor{red}{(???)}$.  Supposing~\eqref{eqn:approximation_error_pf1} and~\eqref{eqn:approximation_error_pf2} hold, and additionally $r < \kappa^{-1/d}$, then some straightforward algebra implies
	\begin{equation*}
	c n r^{d+2} \kappa^{2/d}  \leq \lambda_{\kappa}(\wb{G}_{n,r}) \leq \deg_{\min}(\wb{G}_{n,r}).
	\end{equation*}
	\item \textbf{Deterministic bounds:} 
	Specifically as a consequence of the upper bound  $\lambda_k(\wb{G}_{n,r}) \geq \deg_{\min}(\wb{G}_{n,r})$, we have that
	\begin{equation}
	\label{eqn:approximation_error_pf3}
	\sum_{k = 1}^{\kappa} \bigl(\dotp{v_k}{f_0}_n\bigr)^2 = \sum_{k = 1}^{\kappa} \bigl(\dotp{v_k}{\wb{f}_0}_n\bigr)^2
	\end{equation}
	For any $f \in \Reals^n$, we have
	\begin{equation}
	\norm{f}_n^2 - \frac{f^T \wb{L}_{n,r}^s f}{n \lambda_{\kappa}(\wb{G}_{n,r})}
	\end{equation}

	Neither the equality nor the inequality follow from probabilistic reasoning (except through the reasoning used to establish $\lambda_k(\wb{G}_{n,r}) \geq \deg_{\min}(\wb{G}_{n,r})$). The equality follows from the product graph structure of $\wb{G}_{n,r}$. The inequality is a standard inequality used in the analysis of truncated-series estimators, and would hold for any graph $G$ on the vertices ${\bf X}$.

	\item Under the assumptions~\ref{asmp:smooth_function} and~\ref{asmp:smooth_density}, the graph Sobolev seminorm is upper bounded
	\begin{equation*}
	\wb{f}_0^T \wb{L}_{n,r}^s \wb{f}_0 \leq n^{s + 1}r^{s(d + 2)} B^2
	\end{equation*}
	with probability at least $1 - C_1 r^{-d} \exp\{-c_2 nr^d\}$.
	\item The empirical norm of the histogram estimate $\wb{f}_0$ is lower bounded
	\begin{equation*}
	\norm{\wb{f}_0}_n^2 \geq c \norm{f_0}_{\Leb_2}^2
	\end{equation*}
	with probability at least $1 - \textcolor{red}{???}$.
\end{enumerate}

\subsection{Step 1: Bounds on graph eigenvalues}

\subsection{Step 2: Deterministic bounds}

\subsubsection{Proof of~\eqref{eqn:approximation_error_pf3}.}
The graph $\wb{G}_{n,r}$ is a certain type of product graph, which we term the~\textcolor{red}{(Alden product graph)}. We shall show that all product graphs of this type satisfy an equality similar to~\eqref{eqn:approximation_error_pf3}.

\paragraph{\textcolor{red}{(Alden product graph)}.}
Let $G = ([\mc{M}],E)$ be a graph on $\mc{M} \geq 1$ vertices, and let $n_1, \ldots, n_{\mc{M}}$ each be positive integers; let $N = \sum_{i = 1}^{\mc{M}} n_i$.  The \textcolor{red}{(Alden product)} graph $G^{\boxdot}(G;n_1,\ldots,n_{\mc{M}})$ is defined as
\begin{equation*}
G^{\boxdot}(G;n_1,\ldots,n_{\mc{M}}) = \biggl(\bigcup_{m = 1}^{\mc{M}} \bigcup_{i = 1}^{n_m} (m,i),~~ E^{\boxdot}\biggr),~~\textrm{where}~ \Bigl((\ell,i),(m,j)\Bigr) \in E^{\boxdot}~\textrm{if}~ (\ell,m) \in E.
\end{equation*}
We will simply write $G^{\boxdot}$ when it is clear from context what $G$ and $n_1,\ldots,n_{\mc{M}}$ are. Lemma~\ref{lem:product_eigenvectors} characterizes some of the eigenvectors of $G^{\boxdot}$.
\begin{lemma}
	\label{lem:product_eigenvectors}
	For any $m \in \mc{M}$, and any $i \neq j \in [n_{m}]$, the vector
	\begin{equation}
	\label{eqn:product_eigenvectors}
	g\bigl[(m',k)\bigr] = \1\Bigl\{(m',k) = (m,i)\Bigr\} - \1\Bigl\{(m',k) = (m,j)\Bigr\}
	\end{equation}
	is an eigenvector of $G^{\boxdot}$, with eigenvalue $\deg\Bigl((m,i);G^{\boxdot}\Bigr)$.
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:product_eigenvectors}.]
	content...
\end{proof}
We can see from Lemma~\ref{lem:product_eigenvectors} that any eigenvector $g$ of $G^{\boxdot}$ which does not satisfy~\eqref{eqn:product_eigenvectors} must instead be piecewise constant, i.e. $g\bigl[(m,i)\bigr] = g\bigl[(m,j)\bigr]$ for all $m \in [\mc{M}]$ and $i,j \in [n_m]$. For any such eigenvector, and for any $f: \bigcup_{m = 1}^{\mc{M}} \bigcup_{i = 1}^{n_m} (m,i) \to \Reals$, we have that
\begin{equation*}
\sum_{m = 1}^{\mc{M}} \sum_{i = 1}^{n_m} f\bigl[(m,i)\bigr] g\bigl[(m,i)\bigr] = \sum_{m = 1}^{\mc{M}} \sum_{i = 1}^{n_m} \wb{f}\bigl[(m,i)\bigr] g\bigl[(m,i)\bigr]
\end{equation*}
for $\wb{f}\bigl[(m,i)\bigr] = (n_m)^{-1} \sum_{i = 1}^{n_m} f\bigl[(m,i)\bigr]$. Corollary~\ref{cor:product_eigenvectors} follows immediately from this reasoning.
\begin{corollary}
	\label{cor:product_eigenvectors}
	Let $k$ be an integer such that $\lambda_k(G^{\boxdot}) < \deg\Bigl((m,i);G^{\boxdot}\Bigr)$. Then the eigenvectors $v_1(G^{\boxdot}),\ldots,v_k(G^{\boxdot})$ are all piecewise constant. As a result, for any function $f$,
	\begin{equation*}
	\sum_{k = 1}^{\kappa} \biggl(\sum_{m = 1}^{\mc{M}} \sum_{i = 1}^{n_m} f\bigl[(m,i)\bigr] v_k\bigl[(m,i)\bigr]\biggr)^2 = \sum_{k = 1}^{\kappa} \biggl(\sum_{m = 1}^{\mc{M}} \sum_{i = 1}^{n_m} \wb{f}\bigl[(m,i)\bigr] v_k\bigl[(m,i)\bigr]\biggr)^2
	\end{equation*}
\end{corollary}
We obtain the equality~\eqref{eqn:approximation_error_pf3} by applying Corollary~\ref{cor:product_eigenvectors} to the graph $\wb{G}_{n,r}$.

\subsection{Step 3: Upper bound on the graph Sobolev semi-norm}

\subsection{Step 4: Lower bound on the empirical norm of the histogram estimate}

\end{document}

### 2020-08-19 notes/weekly_notes/7.17.19.NotesForWeek.tex

@@
 	\label{lem:sobolev_1_bound}
-	For any $f \in W^{1,2}(\mathcal{D};1)$, and any $b \geq 1$, we have that there exists a constant $C > 0$ which depends only on $d$ and $p_{\max}$ such that
+	For any $f \in W^{1,2}(\mathcal{D};L)$, and any $b \geq 1$, we have that there exists a constant $c_2 > 0$ which depends only on $d$ and $p_{\max}$ such that
 	\begin{equation}
 	\label{eqn:sobolev_1_bound}
-	\norm{B\beta}_2^2 \leq b C n^2 r^{d + 2}
+	\norm{B\beta}_2^2 \leq L^2 b^2 c_2 n^2 r^{d + 2}
 	\end{equation}
-	with probability at least $1 - \frac{1}{b}$. 
+	with probability at least $1 - \frac{1}{b^2}$. 
 \end{lemma}
@@
 	\end{align*}
-	We will show that for any $f \in \mathcal{W}^{1,2}(\D;1)$, there exists a constant $C$ which depends only on dimension $d$ such that
+	We will show that for any $f \in \mathcal{W}^{1,2}(\D;L)$, there exists a constant $c_4$ which depends only on dimension $d$ such that
 	\begin{equation}
 	\label{eqn:mean_bound}
-	\int_{\D} \int_{\D} (f(x) - f(y))^2\eta_r(x,y) \,dy \,dx \leq C r^{d + 2}
+	\int_{\D} \int_{\D} (f(x) - f(y))^2\eta_r(x,y) \,dy \,dx \leq c_4 L^2 r^{d + 2}
 	\end{equation}

### 2020-08-26 notes/graph_regression2.tex
Updated graph regression document. Results for fixed graph estimation and testing---for Laplacian smoothing and Laplacian eigenmaps---now complete.

@@
 \newcommand{\vardeltapq}{\varDelta^{(p,q)}}
+\newcommand{\lambdavec}{\boldsymbol{\lambda}}
 
@@
 \newcommand{\PLap}{{\bf P}}
+\newcommand{\Id}{I}
 
@@
 	\item For functions $f,g$ which map from $\Xset$ to $\Reals$, we write $\dotp{f}{g}_n := \frac{1}{n}\sum_{i = 1}^{n} f(X_i) g(X_i)$, which we refer to as the \emph{empirical inner product}, and $\norm{f}_n^2 = \dotp{f}{f}_n$, which we refer to as the (squared) \emph{empirical norm}.
+For vectors $v \in \Reals^n$, we write $\norm{v}_p^p := \sum_{i = 1}^{n} \abs{v_i}^p$.
+	\item For a graph $G$, we use $\Lap_G$ for the Laplacian of $G$.
+The matrix $\Lap_G$ is a square, symmetric, positive semi-definite matrix, with $n$ eigenvalues $\lambda_k(G)$ and eigenvector $v_k(G)$ pairs, defined according to
+	\begin{equation*}
+	\Lap_G v_k(G) = \lambda_k(G) v_k(G), ~~\frac{1}{n} \norm{v_k(G)}_2^2 = 1 
+	\end{equation*}
+	and ordered so that $0 \leq \lambda_1(G) \leq \ldots \leq \lambda_n(G)$. For convenience, we write $\Lap_{n,r}$ instead of $\Lap_{G_{n,r}}$. 
 \end{itemize}
@@
 Comment on the different types of relationships between $s$ and $d$.
-(1) Those due to fundamental lower bounds on testing; (2) those due to ``defects'' in the proposed methods, i.e.
+(1) Those due to fundamental upper bounds on testing; (2) those due to ``defects'' in the proposed methods, i.e.
 largeness of the parameter space for Laplacian smoothing, and (3) those (potentially) due to ``defects'' in our proof strategy.
@@
 First, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to functionals of the graph $G$.
-We then analyze the behavior of these functionals with respect to the particular neighborhood graph $G_{n,r}$ and giving high probability (upper or lower )bounds on these functionals.
+We then analyze the behavior of these functionals with respect to the particular neighborhood graph $G_{n,r}$ and give high probability (upper or lower) bounds on these functionals.
 It is in this second step where we invoke our various assumptions on the distribution $P$ and regression function $f_0$.
 
-\section{Fixed graph error bounds}
+\section{Graph-dependent error bounds}
 \label{sec:fixed_graph_error_bounds}
 
-Suppose we observe a graph $G = ([n],E)$, and responses
+Suppose we observe a fixed graph $G = \bigl([n],E\bigr)$ with Laplacian $\Lap_G$ and responses
 \begin{equation}
@@
 \end{equation}
-where $f_0 \in \Reals^n$ and $\varepsilon_i$ are independent $N(0,1)$ noise. The Laplacian eigenmaps estimator of $f_0$ on $G$ is then
+where $f_0 = (f_{0,1},\ldots,f_{0,n}) \in \Reals^n$, and the noise variables $\varepsilon_i$ are independent $N(0,1)$. The Laplacian eigenmaps estimator of $f_0$ on $G$ is then
 \begin{equation}
@@
 \end{equation}
-the Laplacian smoothing estimator of $f_0$ on $G$ is
+where $(\lambda_1(G),v_1(G)),\ldots,(\lambda_n(G),v_n(G))$ are the eigenvalue/eigenvector pairs of $\Lap_G$. The Laplacian smoothing estimator of $f_0$ on $G$ is
 \begin{equation}
 \label{eqn:ls_G}
-\wh{f}_{\LS}(G) = \argmin_{f \in \Reals^n} \biggl\{ \sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_G^{s}  f \biggr\} = (\rho \Lap_G^s + I)^{-1}Y
+\wt{f}_{\LS}(G) := \argmin_{f \in \Reals^n} \biggl\{ \sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_G^{s}  f \biggr\} = (\rho \Lap_G^s + \Id)^{-1}Y.
 \end{equation}
-and their associated test statistics are
+
+In this fixed design context, the hypothesis testing problem becomes
+\begin{equation*}
+\mathbf{H}_0: f_{0} = (0,...,0) ~~\textrm{versus}~~ \mathbf{H}_a: f_{0} \neq (0,...,0)
+\end{equation*}
+and two test statistics for this purpose are
 \begin{equation}
 \label{eqn:le_ts_G}
-T_{\LE}(G) := \Bigl\|\wh{f}_{\LE}(G)\Bigr\|_n^2
+T_{\LE}(G) := \frac{1}{n} \Bigl\|\wh{f}_{\LE}(G)\Bigr\|_2^2 
 \end{equation}
@@
 \label{eqn:ls_ts_G}
-T_{\LS}(G) := \Bigl\|\wt{f}_{\LS}(G)\Bigr\|_n^2
+T_{\LS}(G) := \frac{1}{n} \Bigl\|\wt{f}_{\LS}(G)\Bigr\|_2^2 
 \end{equation}
@@
 
-\subsection{Estimation: Laplacian eigenmaps on a fixed graph}
+\subsection{Error bounds for linear smoothers}
+Let $S \in \Reals^{n \times n}$ be a square, symmetric matrix, and let $\check{f} = SY$ be a linear estimator of $f_0$.
+Both~\eqref{eqn:le_G} and~\eqref{eqn:ls_G} can be written in this form, and Lemma~\ref{lem:linear_smoother_fixed_graph_estimation} is therefore useful in analyzing them.
+Let $\lambdavec(S) = (\lambda_1(S),\ldots,\lambda_n(S)) \in \Reals^n$ denote the eigenvalues of $S$ and let $v_k(S)$ denote the corresponding \emph{unit-norm} eigenvectors, so that $S = \sum_{k = 1}^{n} \lambda_k(S) \cdot v_k(S) v_k(S)^T$.
+Denote $Z_k = v_k(S)^T \varepsilon$, and observe that $Z = (Z_1,\ldots,Z_n) \sim N(0,\Id)$.
+
+\begin{lemma}
+	\label{lem:linear_smoother_fixed_graph_estimation}
+	Let $\check{f} = SY$ for a square, symmetric matrix, $S \in \Reals^{n \times n}$ satisfying $\lambda_n(S) \leq 1$. Then
+	\begin{equation*}
+	\Pbb_{f_0}\biggl(\frac{1}{n}\bigl\|\check{f} - f_0\bigr\|_2^2 \geq \frac{10}{n} \bigl\|\lambdavec(S)\bigr\|_2^2 + \frac{2}{n}\bigl\|(S - I)f_0\bigr\|_2^2\biggr) \leq 1 - \exp\Bigl(-\bigl\|\lambdavec(S)\bigr\|_2^2\Bigr)
+	\end{equation*}
+\end{lemma}
+\begin{proof}[\textbf{Proof of Lemma~\ref{lem:linear_smoother_fixed_graph_estimation}}]
+	It holds that
+	\begin{align*}
+	\frac{1}{n}\bigl\|\check{f} - f_0\bigr\|_2^2 & \leq \frac{2}{n}\Bigl(\bigl\|\check{f} - \Ebb_{f_0}[\check{f}]\bigr\|_2^2 + \bigl\|\Ebb_{f_0}[\check{f}] - f_0\bigr\|_2^2\Bigr) \\ 
+	& = \frac{2}{n}\Bigl(\bigl\|S\varepsilon\bigr\|_2^2 + \bigl\|(S - I)f_0\bigr\|_2^2\Bigr)
+	\end{align*}
+	Writing $\norm{S\varepsilon}_2^2 = \sum_{k = 1}^{n} \lambda_k(S)^2 Z_k^2$, the claim follows from the result of \textcolor{red}{Laurent and Massart} on concentration of $\chi^2$-random variables, which for completeness we restate in Lemma~\ref{lem:chi_square_bound}.
+To be explicit, taking $t = \norm{\lambdavec(S)}_2^2$ in Lemma~\ref{lem:chi_square_bound} completes the proof of Lemma~\ref{lem:le_fixed_graph_estimation}.
+\end{proof}
+
+On the testing side, the statistic $\norm{\check{f}}_2^2 = Y^T S^2 Y$ is a $U$-statistic of order $2$.
+Both~\eqref{eqn:le_ts_G} and~\eqref{eqn:ls_ts_G} can be written in this form, and Lemma~\ref{lem:linear_smoother_fixed_graph_testing} is therefore useful in analyzing them.
+\begin{lemma}
+	\label{lem:linear_smoother_fixed_graph_testing}
+	Let $T = Y^T S^2 Y$ for a square, symmetric matrix $S \in \Reals^{n \times n}$. Define the threshold $t_b$ to be 
+	\begin{equation}
+	t_{b} := \norm{\lambdavec(S)}_2^2 + 2b \norm{\lambdavec(S)}_4^2
+	\end{equation}
+	Suppose the eigenvalues $0 \leq \lambda_{1}(S) \leq \lambda_{n}(S) \leq 1$. Then,
+	\begin{itemize}
+		\item \textbf{Type I error.}
+		\begin{equation}
+		\label{eqn:linear_smoother_fixed_graph_testing_typeI}
+		\Pbb_0\bigl(T > t_b\bigr) \leq \frac{1}{b^2}
+		\end{equation}
+		\item \textbf{Type II error.} Suppose further that
+		\begin{equation}
+		\label{eqn:linear_smoother_fixed_graph_testing_critical_radius}
+		f_0^T S^2 f_0 \geq 4b \norm{\lambdavec(S)}_4^2.
+		\end{equation}
+		Then
+		\begin{equation}
+		\label{eqn:linear_smoother_fixed_graph_testing_typeII}
+		\Pbb_{f_0}\bigl(T \leq t_b\bigr) \leq \frac{4}{b^2} + \frac{8}{b \norm{\lambdavec(S)}_4^{2}} 
+		\end{equation}
+	\end{itemize}
+\end{lemma}
+\begin{proof}[\textbf{Proof of Lemma~\ref{lem:linear_smoother_fixed_graph_testing}}]
+	We compute the mean and variance of $T$ as a function of $f_0$, then apply Chebyshev's inequality.
+
+	\textit{Mean.} Writing $Y = f_0 + \varepsilon$, we make use of the eigendecomposition of $S$ to obtain
+	\begin{equation}
+	\label{pf:linear_smoother_fixed_graph_testing1}
+	\begin{aligned}
+	T & = f_0^T S^2 f_0 + 2 f_0^T S^2 \varepsilon + \varepsilon^T S^2 \varepsilon \\
+	& = f_0^T S^2 f_0 + 2 f_0^T S^2 \varepsilon + \sum_{k = 1}^{n}  \bigl(\lambda_k(S)\bigr)^2 (\varepsilon^T v_k(S))^2 \\
+	& = f_0^T S^2 f_0 + 2 f_0^T S^2 \varepsilon + \sum_{k = 1}^{n}  \bigl(\lambda_k(S)\bigr)^2 Z_k^2,
+	\end{aligned}
+	\end{equation}
+	implying
+	\begin{equation}
+	\label{pf:linear_smoother_fixed_graph_testing_mean}
+	\Ebb_{f_0}[T] = f_0^T S^2 f_0 + \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^2.
+	\end{equation}
+
+	\textit{Variance.} Starting from~\eqref{pf:linear_smoother_fixed_graph_testing1} and recalling the basic fact $\Var(Z_k^2) = 2$, we derive
+	\begin{equation}
+	\label{pf:linear_smoother_fixed_graph_testing_var}
+	\Var_{f_0}[T] \leq 8 f_0^T S^4 f_0 + 4 \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^4 \leq 8 f_0^T S^2 f_0 + 4 \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^2
+	\end{equation}
+	where the second inequality follows since by assumption $\lambda_{n}(S) \leq 1$.
+
+	\textit{Bounding Type I and Type II error.} The bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeI} follows directly from Chebyshev's inequality, along with our above calculations on the mean and variance of $T$.
+
+	The bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeII} also follows from Chebyshev's inequality, as can be seen by the following manipulations,
+	\begin{equation*}
+	\begin{aligned}
+	\Pbb_{f_0}\bigl(T \leq t_b\bigr) & = \Pbb_{f_0}\bigl(T - \Ebb_{f_0}[T] \leq t_b - \Ebb_{f_0}[T]\bigr) \\
+	& \overset{(i)}{\leq} \Pbb_{f_0}\bigl(\abs{T - \Ebb_{f_0}[T]} \geq \abs{t_b - \Ebb_{f_0}[T]}\bigr) \\ 
+	& \overset{(ii)}{\leq} 4 \frac{\Var_{f_0}[T]}{(f_0^T S^2 f_0)^2} \\
+	& \overset{(iii)}{\leq} \frac{32}{f_0^T S^2 f_0} + \frac{4}{b^2} \\
+	& \overset{(iv)}{\leq} \frac{8}{b} \Biggl(\sum_{k = 1}^{n}\bigl(\lambda_k(S)\bigr)^4\Biggr)^{-1/2} + \frac{4}{b^2}
+	\end{aligned}
+	\end{equation*}
+	In the previous expression, $(i)$ and $(ii)$ follow since assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and equation~\eqref{pf:linear_smoother_fixed_graph_testing_mean} together imply $\Ebb_{f_0}(T) - \frac{1}{2}f_0^T S^2f_0 \geq t_b$, $(iii)$ follows from assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and the inequality~\eqref{pf:linear_smoother_fixed_graph_testing_var}, and $(iv)$ follows assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}.
+\end{proof}
+\subsection{Analysis of Laplacian eigenmaps}
 
@@
 	\label{lem:le_fixed_graph_estimation}
-	Let $G$ be a fixed (non-random) graph. Suppose we observe $G$ and responses $Y$ according to the model~\eqref{eqn:fixed_graph_regression_model}. Then for any $1 \leq \kappa \leq n$, 
+	For any integer $1 \leq \kappa \leq n$ and any $s > 0$,
 	\begin{equation}
 	\label{eqn:le_fixed_graph_estimation}
-	\Ebb\Bigl[\norm{\wh{f}_{\LE}(G) - f_0}_n^2\Bigr] \leq 2\biggl(\frac{f_0^T \Lap_{G}^s f_0}{n \bigl(\lambda_{\kappa}(G)\bigr)^s} + \frac{\kappa}{n}\biggr).
-	\end{equation}
-	Additionally, 
-	\begin{equation}
-	\label{eqn:le_fixed_graph_estimation_1}
-	\norm{\wh{f}_{\LE}(G) - f_0}_n^2 \leq 2\biggl(\frac{f_0^T \Lap_{G}^s f_0}{n \bigl(\lambda_{\kappa}(G)\bigr)^s} + \frac{4\kappa}{n}\biggr)
+	\frac{1}{n}\bigl\|\wh{f}_{\LE}(G) - f_0\bigr\|_2^2 \leq 2\frac{f_0^T \Lap_{G}^s f_0}{n\lambda_{\kappa}(G)^s} + \frac{10\kappa}{n}
 	\end{equation}
@@
 \end{lemma}
-The two terms on the right hand side of~\eqref{eqn:le_fixed_graph_estimation} and~\eqref{eqn:le_fixed_graph_estimation_1} represent the squared bias and variance of the estimator $\wh{f}_{\LE}(G)$.
-Note that the squared bias--- and therefore the overall mean squared error---depends on the quadratic form $f_0^T \Lap_{G}^s f_0$ and the eigenvalue $\lambda_{\kappa}(G)$, whereas the variance depends instead on the dimension $\kappa$ of the subspace $V_{\kappa} = \textrm{span}\{v_1,\ldots,v_{\kappa}\}$ to which $\wh{f}_{\LE}(G)$ belongs.
-In this respect, the error bound on the Laplacian eigenmaps estimator resembles the error bound on the \textcolor{red}{orthogonal series} estimator (see e.g.
-\textcolor{red}{Tsybakov}).
+\begin{proof}[\textbf{Proof of Lemma~\ref{lem:le_fixed_graph_estimation}}]
+	Note that if $\lambda_{\kappa}(G) = 0$ the claim is trivial, and we therefore only cover the case when $\lambda_{\kappa}(G) > 0$.
 
-\paragraph{Proof of Lemma~\ref{lem:le_fixed_graph_estimation}:}
+	Letting $\wh{S} = \frac{1}{n}\sum_{k = 1}^{n} v_k(G) v_k(G)^T$, the estimator $\wh{f}_{\LE}(G) = \wh{S}Y$.
+The matrix $\wh{S} \in \Reals^{n \times n}$ is symmetric, and satisfies $\lambda_k(\wh{S}) \in \{0,1\}$ for all $k \in [n]$.
+The claim then follows from Lemma~\ref{lem:le_fixed_graph_estimation} upon noting first that $\norm{\lambdavec(\wh{S})}_2^2 = \kappa$, and second that
+	\begin{equation}
+	\label{pf:le_fixed_graph_estimation_1}
+	\begin{aligned}
+	\bigl\|(\wh{S} - I)f_0\bigr\|_2^2 & = \sum_{k = \kappa + 1}^{n} \bigl(f_0^T v_k(G)\bigr)^2 \\
+	& \leq \frac{1}{\lambda_{\kappa}(G)^s}\sum_{k = \kappa + 1}^{n} \lambda_{k}(G)^s \bigl(f_0^T v_k(G)\bigr)^2 \\
+	& \leq \frac{1}{\lambda_{\kappa}(G)^s}\sum_{k = 1}^{n} \lambda_{k}(G)^s \bigl(f_0^T v_k(G)\bigr)^2 \\
+	& = \frac{f_0^T \Lap_G^s f_0}{\lambda_{\kappa}(G)^s}.
+	\end{aligned}
+	\end{equation} 
+\end{proof}
 
-\textcolor{red}{(TODO)}
-
-\subsection{Estimation: Laplacian smoothing on a fixed graph.}
-
-\begin{lemma}
-	\label{lem:ls_fixed_graph_estimation}
-	Suppose we observe a graph $G$ and responses $Y$ according to the model~\eqref{eqn:fixed_graph_regression_model}. Then for any $\rho > 0$ and positive integer $s$,
-	\begin{equation}
-	\label{eqn:ls_fixed_graph_estimation}
-	\Ebb\Bigl[\norm{\wt{f}_{\LS}(G) - f_0}_n^2\Bigr] \leq 2\biggl(\frac{\rho \cdot f_0^T \Lap_{G}^s f_0}{n} + \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{(\rho \cdot \lambda_{k}(G)^s + 1)^2}\biggr)
-	\end{equation}
-	Additionally,
-	\begin{equation}
-	\label{eqn:ls_fixed_graph_estimation_prob}
-	\Pbb\Biggl[\norm{\wt{f}_{\LS}(G) - f_0}_n^2 \leq 2\biggl(\frac{\rho \cdot f_0^T \Lap_{G}^s f_0}{n} + \frac{4}{n}\sum_{k = 1}^{n} \frac{1}{(\rho \cdot \lambda_{k}(G)^s + 1)^2}\biggr)\Biggr] \geq 1 - \exp\biggl(-\sum_{k = 1}^{n}\frac{1}{(\rho \lambda_k(G)^s + 1)^2}\biggr)
-	\end{equation}
-\end{lemma}
-
-\paragraph{Proof of Lemma~\ref{lem:ls_fixed_graph_estimation}:}
-
-\textcolor{red}{(TODO)}
-
-\subsection{Testing: Laplacian eigenmaps on a fixed graph.}
+In Lemma~\ref{lem:le_fixed_graph_testing}, we upper bound the testing error of a test based on $T_{\LE}(G)$.
 
@@
 	\label{lem:le_fixed_graph_testing}
-	Suppose we observe a graph $G$ and responses $Y$ according to the model~\eqref{eqn:fixed_graph_regression_model}. Then for any $1 \leq \kappa \leq n$, each of the following holds for any $b \geq 1$,
+	Define the threshold $\wh{t}_b$ to be
+	\begin{equation*}
+	\wh{t}_b := \frac{\kappa}{n} + \frac{2b\sqrt{\kappa}}{n}
+	\end{equation*}
+	Then each of the following hold for any $\kappa \in [n]$ and any $b \geq 1$,
 	\begin{enumerate}
@@
 		\label{eqn:graph_spectral_type_I_error}
-		\Pbb\biggl(T_{\LE}(G) \geq \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}\biggr) \leq \frac{2}{b^2}.
+		\Pbb_{f_0}\biggl(T_{\LE}(G) > \wh{t}_b\biggr) \leq \frac{1}{b^2}.
 		\end{equation}
-		\item \textbf{Type II error:} Under the alternative hypothesis $f_0 \neq 0$, suppose that
+		\item \textbf{Type II error:} Suppose that
 		\begin{equation}
-		\label{eqn:fixed_graph_testing_critical_radius_1}
-		\Bigl\|\sum_{k = 1}^{\kappa} \dotp{f_0}{v_k}_n v_k\Bigr\|_n^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}}.
+		\label{eqn:le_fixed_graph_testing_critical_radius}
+		\frac{1}{n} \norm{f_0}_2^2 \geq \frac{f_0^T \Lap_G^s f_0}{n \lambda_{\kappa}(G)^s} + \frac{4b\sqrt{\kappa}}{n}
 		\end{equation}
@@
 		\label{eqn:graph_spectral_type_II_error}
-		\Pbb\biggl(T_{\LE}(G) \leq \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}\biggr) \leq \frac{2}{b^2} + \frac{8}{b\sqrt{2\kappa}}.
+		\Pbb_{f_0}\biggl(T_{\LE}(G) \leq \wh{t}_b\biggr) \leq \frac{4}{b^2} + \frac{8}{b\sqrt{\kappa}}.
 		\end{equation}
-		In particular if
-		\begin{equation}
-		\label{eqn:fixed_graph_testing_critical_radius}
-		\frac{1}{n} \sum_{i = 1}^{n} (f_{0,i})^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{f_0^T L_G^s f_0}{n\lambda_{\kappa}(G)^s}
-		\end{equation}
-		then~\eqref{eqn:fixed_graph_testing_critical_radius_1} and thus~\eqref{eqn:graph_spectral_type_II_error} follow.
 	\end{enumerate}
 \end{lemma}
+\begin{proof}[\textbf{Proof of Lemma~\ref{lem:le_fixed_graph_testing}}]
+	Letting $\wh{S} = \frac{1}{n}\sum_{k = 1}^{\kappa} v_k(G) v_k(G)^T$, our test statistic $T_{\LE}(G) = \frac{1}{n} Y^T \wh{S}^2 Y$.
+The matrix $\wh{S}$ is symmetric and positive semidefinite, with eigenvalues $\lambda_k(\wh{S}) \in \{0,1\}$ for all $k \in [n]$, so $\norm{\lambdavec(\wh{S})}_2^2 = \norm{\lambdavec(\wh{S})}_4^4 = \kappa$ and clearly $n \wh{t}_b = \norm{\lambdavec(\wh{S})}_2^2 + 2b \norm{\lambdavec(\wh{S})}_4^2$.
+The desired result thus follows from Lemma~\ref{lem:linear_smoother_fixed_graph_testing}.
+To see that the conditions of this Lemma are met, note first that $\lambda_k(\wh{S}) \in \{0,1\}$ for all $k \in [n]$, and so it remains only to verify \eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}.
+Invoking assumption~\eqref{eqn:le_fixed_graph_testing_critical_radius} along with the inequality~\eqref{pf:le_fixed_graph_estimation_1}, we obtain
+	\begin{align*}
+	f_0^T \wh{S}^2 f_0 & = \norm{f_0}_2^2 - \sum_{k = \kappa + 1}^{n} \bigl(f_0^T v_k(G)\bigr)^2 \\
+	& \geq \frac{f_0^T \Lap_G^s f_0}{\lambda_{\kappa}(G)^s} + 4b\sqrt{\kappa} - \sum_{k = \kappa + 1}^{n} \bigl(f_0^T v_k(G)\bigr)^2 \\
+	& \geq 4b\sqrt{\kappa}.
+	\end{align*}
+	Reiterating that $\norm{f_0}_4^2 = \sqrt{\kappa}$, we see that~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} is met, completing the proof.
+\end{proof}
 
-\subsection{Testing: Laplacian smoothing on a fixed graph.}
+\subsection{Analysis of Laplacian Smoothing}
+In Lemma~\ref{lem:ls_fixed_graph_estimation}, we upper bound the mean squared error of $\wt{f}_{\LS}(G)$.
+\begin{lemma}
+	\label{lem:ls_fixed_graph_estimation}
+	For any $\rho,s > 0$,
+	\begin{equation}
+	\label{eqn:ls_fixed_graph_estimation_prob}
+	\frac{1}{n}\bigl\|\wt{f}_{\LS}(G) - f_0\bigr\|_2^2 \leq \frac{2\rho}{n} \bigl(f_0^T \Lap_{G}^s f_0\bigr) + \frac{10}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k}(G)^s + 1\bigr)^2}
+	\end{equation}
+	with probability at least $1 - \exp\Bigl(-\sum_{k = 1}^{n}\bigl(\rho \lambda_k(G)^s + 1\bigr)^{-2}\Bigr)$.
+\end{lemma}
+\begin{proof}[\textbf{Proof of Lemma~\ref{lem:ls_fixed_graph_estimation}}]
+	Letting $\wt{S} = (\Id + \rho \Lap_G^s)^{-1}$, the estimator $\wt{f}_{\LS}(G) = \wt{S}Y$.
+The matrix $\wt{S} \in \Reals^{n \times n}$ is symmetric, and satisfies $\lambda_k(\wt{S}) \in (0,1)$ for all $k \in [n]$.
+The claim then follows from Lemma~\ref{lem:linear_smoother_fixed_graph_estimation} upon noting first that
+	\begin{equation*}
+	\bigl\|\lambdavec(\wt{S})\bigr\|_2^2 = \sum_{k = 1}^{n} \frac{1}{\bigl(1 + \rho \lambda_k(G)^s\bigr)^2}
+	\end{equation*} 
+	and second that
+	\begin{equation*}
+	\begin{aligned}
+	\bigl\|(\wt{S} - I)f_0\bigr\|_2^2 & = f_0^T \Lap_G^{s/2} \Lap_G^{-s/2}\bigl(\wt{S} - \Id\bigr) \Lap_G^{-s/2} \Lap_G^{s/2} f_0 \\
+	& \leq f_0^T \Lap_G^s f_0 \cdot \lambda_n\Bigl(\Lap_G^{-s/2}\bigl(\wt{S} - \Id\bigr)\Lap_G^{-s/2}\Bigr) \\
+	& = f_0^T \Lap_G^s f_0 \cdot \max_{k \in [n]} \biggl\{\frac{1}{\lambda_k(G)^s} \Bigl(1 - \frac{1}{\rho\lambda_k(G)^s + 1}\Bigr) \biggr\} \\
+	& = f_0^T \Lap_G^s f_0 \cdot \rho.
+	\end{aligned}
+	\end{equation*} 
+	where we write $\Lap_G^{-1}$ for the pseudoinverse of $\Lap_G$.
+\end{proof}
+
+In Lemma~\ref{lem:ls_fixed_graph_testing}., we upper bound the testing error of a test based on $T_{\LS}(G)$.
+\begin{lemma}
+	\label{lem:ls_fixed_graph_testing}
+	Define the threshold $\wt{t}_b$ to be
+	\begin{equation*}
+	\wt{t}_b := \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k(G)^s + 1\bigr)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k(G)^s + 1\bigr)^4}}
+	\end{equation*}
+	Then each of the following holds for any $\rho,s > 0$ and any $b \geq 1$.
+	\begin{itemize}
+		\item \textbf{Type I error.}
+		\begin{equation}
+		\label{eqn:ls_fixed_graph_testing_typeI}
+		\Pbb_0\Bigl(T_{\LS}(G) > \wt{t}_b\Bigr) \leq \frac{1}{b^2}
+		\end{equation}
+		\item \textbf{Type II error.} If
+		\begin{equation}
+		\label{eqn:ls_fixed_graph_testing_critical_radius}
+		\frac{1}{n}\norm{f_0}_2^2 \geq \frac{2 \rho}{n} \bigl(f_0^T \Lap_G^s f_0\bigr) + \frac{4b}{n} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G)^s + 1)^4} \Biggr)^{1/2}
+		\end{equation}
+		then
+		\begin{equation}
+		\label{eqn:ls_fixed_graph_testing_typeII}
+		\Pbb_{f_0}\Bigl(T_{\LS}(G) \leq \wt{t}_b\Bigr) \leq \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G)^s + 1)^4} \Biggr)^{-1/2}
+		\end{equation}
+	\end{itemize}
+\end{lemma}
+
+\begin{proof}[\textbf{Proof of Lemma~\ref{lem:ls_fixed_graph_testing}}]
+	Let $\wt{S} := (\rho \Lap^s + \Id)^{-1}$.
+The matrix $\wt{S} \in \Reals^{n \times n}$ is symmetric and positive semidefinite, and our test statistic $T_{\LS}(G) = \frac{1}{n}Y^T \wt{S}^2 Y$.
+The desired result thus follows from Lemma~\ref{lem:linear_smoother_fixed_graph_testing}.
+To see that the conditions of this Lemma are satisfied, we first note that since
+	\begin{equation*}
+	\lambda_k(\wt{S}) = \frac{1}{(\rho\lambda_k(G)^s + 1)}
+	\end{equation*}
+	and $\rho, \lambda_k(G) > 0$, it is evident that $\lambda_{n}(\wt{S}) \leq 1$.  Then, by assumption~\eqref{eqn:ls_fixed_graph_testing_critical_radius}
+	\begin{equation*}
+	f_0^{T} \wt{S}^2 f_0 = \norm{f_0}_2^2 - f_0^T(I - \wt{S}^2)f_0 \geq 2 \rho (f_0^T \Lap^s f_0) + f_0^T(I - \wt{S}^2)f_0 + 4b \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G)^s + 1)^4} \Biggr)^{-1/2},
+	\end{equation*}
+	and along with the following calculations,
+	\begin{equation*}
+	\begin{aligned}
+	f_0^T \Bigl(\Id - \wt{S}^2\Bigr) f_0  & = f_0^T L^{s/2} L^{-s/2}\Bigl(\Id - \wt{S}^2\Bigr) L^{-s/2} L^{s/2} f_0 \\ 
+	& \leq f_0^T L^{s} f_0 \cdot  \lambda_{\max}\biggl(L^{-s/2}\Bigl(\Id - \wt{S}^2\Bigr) L^{-s/2}\biggr) \\ 
+	& \overset{(i)}{=}  f_0^T L^{s} f_0 \cdot \max_{k} \biggl\{ \frac{1}{\lambda_k(G)^s} \Bigl(1 - \frac{1}{(\rho \lambda_k(G)^s + 1)^2}\Bigr) \biggr\} \\
+	& \overset{(ii)}{\leq} f_0^T L^{s} f_0 \cdot 2\rho,
+	\end{aligned}
+	\end{equation*}
+	we have that
+	\begin{equation*}
+	f_0^{T} \wt{S}^2 f_0 \geq 2b \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G)^s + 1)^4} \biggr)^{-1/2}.
+	\end{equation*} 
+	In other words condition~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} in Lemma~\ref{lem:linear_smoother_fixed_graph_testing} is met, and applying that Lemma completes the proof.
+
+	(In the previous derivation: in $(i)$ the maximum is over all indices $k$ such that the eigenvalue $\lambda_k(G)$ is strictly positive; and $(ii)$ follows from the basic algebraic identity $1 - 1/(1 + \rho x)^2 \leq 2 \rho x$ for any $x, \rho > 0$.
+\end{proof}
 
@@
 to the random geometric graph $G_{n,r}$.
-To do so, we must establish the following (high probability) bounds: (i) an upper bound on the graph Sobolev seminorm of the regression function $f_0 \Lap_{n,r}^s f_0$, (ii) and lower boud on the Laplacian eigenvalue $\lambda_{\kappa}(G_{n,r})$ as a function of $\kappa$, and (iii) a lower bound on the empirical norm $\norm{f}_{n}$ as a function of $\norm{f}_{L^2(P)}$.
+To do so, we must establish the following (high probability) bounds: (i) an upper bound on the graph Sobolev seminorm of the regression function $f_0 \Lap_{n,r}^s f_0$, (ii) a lower bound on the Laplacian eigenvalue $\lambda_{\kappa}(G_{n,r})$ as a function of $\kappa$, and (iii) a lower bound on the empirical norm $\norm{f}_{n}$ as a function of $\norm{f}_{L^2(P)}$.
 
@@
 \section{Concentration Inequalities}
+\begin{lemma}
+	\label{lem:chi_square_bound}
+	Let $\xi_1,\ldots,\xi_N$ be independent $N(0,1)$ random variables, and let $U := \sum_{k = 1}^{N} a_k(\xi_k^2 - 1)$.  Then for any $t > 0$,
+	\begin{equation*}
+	\Pbb\Bigl[U \geq 2 \norm{a}_2 \sqrt{t} + 2 \norm{a}_{\infty}t\Bigr] \leq \exp(-t).
+	\end{equation*}
+	In particular if $a_k = 1$ for each $k = 1,\ldots,N$, then
+	\begin{equation*}
+	\Pbb\Bigl[U - N \geq 2\sqrt{N t} + 2t\Bigr] \leq \exp(-t).
+	\end{equation*}
+\end{lemma}
 

### 2020-08-28 notes/weekly_notes/8.26.20.NotesForWeek.tex
Added a sharper analysis of the expected weighted difference operator, relating it to the density-weighted Laplace operator.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{tikz}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\iid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}
\newcommand{\Lap}{L}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}
\newcommand{\Id}{I}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Vset}{\mathcal{V}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 8/26/2020 - 9/2/2020}
\author{Alden Green}
\date{\today}
\maketitle

In this week's notes, I state and prove all results regarding graph Sobolev semi-norms that we need for our graph regression work.

\section{Bounds on the graph Sobolev semi-norm}
In this section we will state and prove three different Lemmas involving the graph Sobolev semi-norm $f_0 \Lap_{n,r}^s f_0$. In each Lemma, we will derive that the asymptotic relation
\begin{equation*}
f_0^T \Lap_{n,r}^s f_0 \lesssim n \bigl(nr^{d + 2}\bigr)^s
\end{equation*}
holds with high probability, whenever $r \to 0$ sufficiently slowly as $n \to \infty$. The difference between the three Lemmas will be the assumptions we make, and the resulting permissible range for $r$. Throughout, we will make heavy use of Taylor's Theorem, which we briefly recall. For an open set $U \subseteq \Reals^d$ and a function $g \in C^{\infty}(U)$, it holds that for all $x,x' \in U$, and for every $s \geq 1$,
\begin{equation*}
g(x') = \sum_{\abs{\alpha} = 0}^{s - 1} g^{(\alpha)}(x) \cdot (x' - x)^{\alpha}  + R_g^{(s)}(x',x)
\end{equation*}
where for concision we have written 
\begin{equation*}
g^{(\alpha)}(x) = \frac{1}{\abs{\alpha}!} \bigl(D^{\alpha}g\bigr)(x),~~ R_g^{(s)}(x',x) = \frac{1}{s!} \sum_{\abs{\alpha} = s} (x' - x)^{\alpha} \int_{0}^{1} (1 - t)^{\abs{\alpha}}  g^{(\alpha)}\bigl(x + t(x' - x)\bigr) \,dt
\end{equation*}
and $x^{\alpha} = \prod_{i = 1}^{d} x_i^{\alpha_i}$ for $x,\alpha \in \Reals^d$.

\subsection{First-order graph Sobolev semi-norm.}
We begin by upper bounding $f_0^T \Lap_{n,r} f_0$ under the assumption $f_0 \in H^1(\Xset,M)$. This will be the easiest of the three Lemmas to prove, while still giving some sense of how the proofs develop in the more complicated cases. The upper bound will depend on the kernel $K$; we assume throughout that $K:[0,\infty) \to [0,\infty)$ has support on $[0,1]$ with
\begin{equation*}
\int_{\Reals^d} K(\norm{z}) \,dz = 1,
\end{equation*}
and we let
\begin{equation*}
I_{\alpha} := \int_{B(0,1)} (z)^{\alpha} K\bigl(\norm{z}\bigr) \,dz
\end{equation*}
where $B(0,1) \subseteq \Reals^d$ is the unit ball centered at the origin.
\begin{lemma}
	\label{lem:first_order_graph_sobolev_seminorm}
	Suppose $\Xset$ is a Lipschitz domain, and that $p(x) \leq p_{\max}$ for all $x \in \Xset$. Then, for any $f \in H^1(\Xset,M)$, there exists a constant $C > 0$ which may depend only on $\Xset$ and $d$ such that
	\begin{equation}
	f^T \Lap_{n,r} f \leq \frac{C}{\delta} n^2 r^{d + 2} M^2
	\end{equation}
	with probability at least $1 - \delta$.
\end{lemma}
\begin{proof}
	We will show that
	\begin{equation*}
	\Ebb\Bigl[f^T \Lap_{n,r} f\Bigr] \leq C n^2 r^{d + 2} M^2,
	\end{equation*}
	whence the claim follows immediately by Markov's inequality (recall that $\Lap_{n,r}$ is positive semi-definite, and therefore $f^T \Lap_{n,r} f$ is a non-negative random variable).

	Since
	\begin{equation*}
	f^T \Lap_{n,r} f = \frac{1}{2}\sum_{i, j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 A_{ij},
	\end{equation*}
	it follows that
	\begin{equation}
	\label{pf:first_order_graph_sobolev_seminorm_1}
	\Ebb\Bigl[f^T \Lap_{n,r} f\Bigr] = \frac{n(n - 1)}{2} \Ebb\biggl[\Bigl(f(X') - f(X)\Bigr)^2 K_r(X',X)\biggr]
	\end{equation}
	where $X$ and $X'$ are random variables independently drawn from $P$.

	For the remainder of this proof, we will assume that $f \in C^{\infty}(\Xset)$, which we may do without loss of generality because $C^{\infty}(\Xset)$ is dense in $H^1(\Xset)$. Moreover, since $\Xset$ is a Lipschitz domain there exists an extension $g \in C^{\infty}(\Rd)$ of $f$ satisfying
	\begin{equation}
	\label{pf:first_order_graph_sobolev_seminorm_2}
	|g|_{H^1(\Rd)} \leq C|f|_{H^1(\Xset)} \leq C M,
	\end{equation}
	for a constant $C$ which may depend on $\Xset$ and $d$, but does not depend on $M$ or $f$. Since $g(x) = f(x)$ and $p(x) \leq p_{\max}$ for all $x \in \Xset$, 
	\begin{align}
	\Ebb\Bigl[\bigl(f(X') - f(X)\bigr)^2K_r(X',X)\Bigr] & = \Ebb\Bigl[\bigl(g(X') - g(X)\bigr)^2K_r(X',X)\Bigr] \nonumber \\
	& \leq p_{\max}^2 \int_{\Reals^d} \int_{\Reals^d} \bigl(g(x') - g(x)\bigr)^2 K_r(x',x) \,dx' \,dx. \label{pf:first_order_graph_sobolev_seminorm_3}
	\end{align}
	It remains only to bound the integral. Taking a first-order Taylor expansion of $g$, we get
	\begin{align*}
	\int_{\Reals^d} \int_{\Reals^d} \bigl(g(x') - g(x)\bigr)^2 K_r(x',x) \,dx \,dx' & = \int_{\Rd} \int_{\Rd} \biggl\{\sum_{\abs{\alpha} = 1}G_{\alpha}(x,x') (x' - x)^{\alpha}\biggr\}^2 K_r(x',x) \,dx' \,dx \\ 
	& \overset{(i)}{\leq}  \int_{\Rd} \int_{\Rd} \biggl\{\sum_{\abs{\alpha} = 1}\Bigl[G_{\alpha}(x,x')\Bigr]^2\biggr\} \norm{x' - x}^2  K_r(x',x) \,dx' \,dx \\
	& \overset{(ii)}{\leq} \int_{\Rd} \int_{\Rd} \biggl\{\int_{0}^{1} \sum_{\abs{\alpha} = 1} \Bigl[g^{(\alpha)}\bigl(x + t(x' - x)\bigr)\Bigr]^2 \,dt \biggr\} \norm{x' - x}^2 K_r(x',x) \,dx' \,dx \\
	& \overset{(iii)}{=} r^{2 + d} \int_{\Rd} \int_{B(0,1)} \biggl\{\int_{0}^{1} \sum_{\abs{\alpha} = 1} \Bigl[g^{(\alpha)}\bigl(x + trz\bigr)\Bigr]^2 \,dt\biggr\} \norm{z}^2 K\bigl(\norm{z}\bigr)  \,dz \,dx \\
	& \overset{(v)}{=} r^{2 + d} \int_{B(0,1)} \int_{0}^{1} \biggl\{\int_{\Rd} \sum_{\abs{\alpha} = 1} \Bigl[g^{(\alpha)}\bigl(x + trz\bigr)\Bigr]^2 \,dx \biggr\} \norm{z}^2 K\bigl(\norm{z}\bigr) \,dt \,dz \\
	& \leq r^{2 + d} |g|_{H^1(\Rd)}.
	\end{align*}
	In the above manipulations, $(i)$ follows by the Cauchy-Schwarz inequality, $(ii)$ follows by Jensen's inequality, $(iii)$ follows from the change of variables $z = (x' - x)/r$, and $(v)$ follows by Fubini's Theorem. Along with~\eqref{pf:first_order_graph_sobolev_seminorm_1}-\eqref{pf:first_order_graph_sobolev_seminorm_3}, this proves Lemma~\ref{lem:first_order_graph_sobolev_seminorm}.
\end{proof}

A few remarks:
\begin{itemize}
	\item \textbf{Bounded norm vs. bounded semi-norm.}
	\item \textbf{Markov's inequality, as opposed to exponential concentration.}
	\item \textbf{Simpler proof and stronger concentration under Holder assumption.}
\end{itemize}

\subsection{General graph Sobolev semi-norm}

When $s > 1$, in order to obtain the desired upper bound on the graph Sobolev semi-norm we must make some additional assumptions regarding the behavior of $f_0$ near the boundary of $\Xset$, as well as on the smoothness of the density function $p$.
\begin{theorem}
	\label{thm:graph_sobolev_seminorm}
	Suppose $\Xset $ is a Lipschitz domain, and that the density $p \in C^{s - 1}(\Xset,p_{\max})$. Then, for any $f \in H_0^s(\Xset,M)$ there exists a constant $C > 0$ which may depend only on $X$ and $d$ such that
	\begin{equation}
	\label{eqn:graph_sobolev_seminorm}
	f^T \Lap_{n,r}^s f \leq \frac{C}{\delta} n^{s + 1} r^{s(d + 2)}M^2
	\end{equation}
	with probability at least $1 - \delta$.
\end{theorem}
The proof of Lemma~\ref{thm:graph_sobolev_seminorm} will be considerably longer than that of Lemma~\ref{lem:first_order_graph_sobolev_seminorm}. In anticipation of this proof, it will be helpful to introduce some notation. For $j \in [n]$ and a function $f: \Xset \to \Reals$, define the weighted difference operator $D_jf: \Xset \to \Reals$ to be
\begin{equation*}
\bigl(D_jf\bigr)(x) = \bigl(f(X_j) - f(x)\bigr) \frac{K_r(X_j,x)}{r^d},
\end{equation*}
and for $k \in [n]^q$ for some $q \in \mathbb{N}$, define the iterated difference operator $D_{jk}f: \Xset \to \Reals$ via the recursive formula
\begin{equation*}
\bigl(D_{jk}f\bigr)(x) = \bigl(D_j \circ D_kf\bigr)(x)
\end{equation*}
The matrix $\Lap_{n,r} \in \Reals^{n \times n}$ can then be extended to an operator over $C^{\infty}(\Xset)$, which in an abuse of notation\footnote{We repeat notation because it results in \eqref{eqn:graph_Laplacian_out_of_sample} agreeing with \textcolor{red}{(2)} in-sample, i.e. $\bigl(\Lap_{n,r}^qf\bigr)(X_i) = (\Lap_{n,r}^qf)_i$ for all $i \in [n]$.} we also denote by $\Lap_{n,r}$,
\begin{equation*}
\frac{1}{r^d}\bigl(\Lap_{n,r}f\bigr)(x) := -\sum_{j = 1}^{n} \bigl(D_jf\bigr)(x)
\end{equation*}
so that
\begin{equation}
\label{eqn:graph_Laplacian_out_of_sample}
\frac{1}{r^{dq}}\bigl(\Lap_{n,r}^qf\bigr)(x) = (-1)^q \sum_{k \in [n]^q} \bigl(D_kf\bigr)(x).
\end{equation} 
Consequently, a few standard computations show that the graph Sobolev semi-norm can be written as 
\begin{equation*}
f^T \Lap_{n,r}^s f = 
\begin{dcases}
\sum_{i = 1}^{n} \sum_{k \in [n]^q} \sum_{\ell \in [n]^q} \bigl(D_kf\bigr)(X_i) \cdot \bigl(D_\ell f\bigr)(X_i),~~ & \textrm{if $s$ is even, for $q = s/2$} \\
\sum_{i,j = 1}^{n} \sum_{k \in [n]^q} \sum_{\ell \in [n]^q} \bigl(d_jD_kf\bigr)(X_i) \cdot \bigl(d_jD_{\ell}f\bigr)(X_i) \cdot K_r(X_i,X_j),~~ & \textrm{if $s$ is odd, for $q = (s - 1)/2$}
\end{dcases}
\end{equation*}
where $\bigl(d_jf\bigr)(x) := f(X_j) - f(x)$ is the simple difference operator.

In order to prove~\eqref{eqn:graph_sobolev_seminorm}, we will proceed as in Lemma~\ref{lem:first_order_graph_sobolev_seminorm} by first upper bounding $\Ebb[f^T \Lap_{n,r}^s f]$ and then applying Markov's inequality. Obtaining the desired upper bound on $\Ebb[f^T \Lap_{n,r}^s f]$ is split into \textcolor{red}{N} Lemmas, Lemmas \textcolor{red}{A - B}. Lemma~\ref{lem:expected_difference_operator} controls the expectation of the ``leading terms'' $\Ebb[D_kf(x)]$, when $k \in (n)^q$ and $x$ is sufficiently in the interior of $\Xset$. 
\begin{lemma}
	\label{lem:expected_difference_operator}
\end{lemma}

\section{Review}
In this section, I put results that will (probably) not be need for our graph regression paper, but which are still nice to know.

Let $\Delta_r$ be the expected weighted difference operator
\begin{equation}
\Delta_rf(x) := \Ebb\Bigl[\bigl(D_if\bigr)(x)\Bigr] = \int_{\Xset} \bigl(f(x') - f(x)\bigr) K_r(x',x) p(x') \,dx
\end{equation}
and define $\Delta_r^q = \Delta_r \circ \cdots \circ \Delta_r$, where the operator is composed a total of $q$ times. 
\begin{lemma}
	\label{lem:expected_difference_holder}
	Let $q \geq 1$ and $s \geq 2q + 1$ be integers. The iterated expected weighted difference operator  $\Delta_r^qf(x) = \Ebb[D_kf(x)]$ for any $k \in (n)^q$. If $f \in C^s(\Xset,M)$ and $p \in C^{s-1}(\Xset,M)$, then for any $x \in \Xset$ such that $B(x,qr) \subset \Xset$, it holds that
	\begin{equation*}
	\Bigl\|\bigl(\Delta_r^qf\bigr)(x) - r^{2q} \sigma_K^{2q} \bigl(\Delta_{p}^qf\bigr)(x)\Bigr\| \leq C \cdot
	\begin{dcases}
	r^{2q + 1},& ~~\textrm{if $s = 2q + 1$} \\
	r^{2q + 2},& ~~\textrm{if $s = 2q + 2$.}
	\end{dcases}
	\end{equation*}
	Recall that $\Delta_pf = -\frac{1}{2p}\mathrm{div}(p^2 \nabla f)$ is the density weighted Laplace operator.
\end{lemma}
\begin{proof}
	We will actually prove a more general claim. Let $f$ be a function defined on the domain $\Xset$, which additionally satisfies $f \in C^s(\Vset,M)$ for some open set $\Vset \subset \Xset$\footnote{Technically speaking, we mean that the restriction $\restr{f}{\Vset}$ of $f$ to $\Vset$ satisfies $\restr{f}{\Vset} \in C^s(\Vset,M)$, but we dispense with this technicality for notational simplicity.}, and $s \geq 2q + 1$. For any $x \in V$ such that $B(x,qr) \subset V$, we will show
	\begin{equation}
	\label{eqn:expected_difference_holder_0}
	\Delta_r^qf(x) = r^{2q} I_{2e_1}^q \Delta_p^qf(x) + \sum_{t = q + 1}^{\floor{(s - 1)/2}}r^{2t} f_{2t}(x) + O\bigl(M p_{\max}r^s\bigr)
	\end{equation} 
	for functions $f_u$ which satisfy $f_u \in C^{s - u}(\Vset_{qr},M)$. To see that this implies the desired claim, simply take $\Vset = \Xset$ and note that when $s = 2q + 1$ or $s = 2q + 2$, the sum on the right hand side of~\eqref{eqn:expected_difference_holder_0} is empty.

	We prove~\eqref{eqn:expected_difference_holder_0} by induction on $q$. In this proof, the functions $f_u$ may differ from line to line, but they will always satisfy $f_u \in C^{s - u}(\Vset_{qr},M)$. When $q = 1$, the expected difference operator
	\begin{equation*}
	\Delta_r^qf(x) = \Delta_rf(x) = -\frac{1}{r^d}\int_{\Xset} \bigl(f(x') - f(x)\bigr) K_r(x',x) p(x') \,dx' =  -\frac{1}{r^d}\int_{\Vset} \bigl(f(x') - f(x)\bigr) K_r(x',x) p(x') \,dx' 
	\end{equation*}
	where we may restrict the integral to $\Vset$ since $B(x,r) \subset \Vset$, and $K$ has support on $[0,1]$. Then, taking Taylor expansions of $f(x')$ and $p(x')$ around $x' = x$ gives
	\begin{equation}
	\begin{aligned}
	\label{eqn:expected_difference_holder_1}
	\Delta_rf(x) & = -\frac{1}{r^d} \sum_{\abs{\alpha} = 1}^{s - 1} \sum_{\abs{\beta} = 0}^{s - 2}  f^{(\alpha)}(x) p^{(\beta)}(x) \int_{\Vset} (x' - x)^{\alpha + \beta} K_r(x',x) \,dx' + O\bigl(M p_{\max}r^s\bigr) \\
	& = -\sum_{\abs{\alpha} = 1}^{s - 1} \sum_{\abs{\beta} = 0}^{s - 2}  r^{\abs{\alpha + \beta}} f^{(\alpha)}(x) p^{(\beta)}(x) \int_{B(0,1)} (z)^{\alpha + \beta} K\bigl(\norm{z}\bigr) \,dz + O\bigl(M p_{\max}r^s\bigr) \\
	& = -\sum_{\abs{\alpha} = 1}^{s - 1} \sum_{\abs{\beta} = 0}^{s - 2}  r^{\abs{\alpha + \beta}} f^{(\alpha)}(x) p^{(\beta)}(x) I_{\alpha + \beta} + O\bigl(M p_{\max}r^s\bigr) \\
	& =: -\sum_{u = 1}^{s - 1} f_u(x) r^u + O\bigl(M p_{\max}r^s\bigr)
	\end{aligned}
	\end{equation}
	where in the first equality we use the bounds $|R_f^{(s)}(x,x')| \leq M\norm{x - x'}^s$ and $|R_p^{(s - 1)}(x,x')| \leq p_{\max}\norm{x - x'}^{s - 1}$ to control the remainder term, the second equality follows since $B(x,r) \subseteq \Vset$, and the final equality follows by putting
	\begin{equation}
	\label{eqn:expected_difference_holder_2}
	f_u(x) = \sum_{\abs{\alpha} = 1} \sum_{\abs{\beta} = u - \abs{\alpha}} I_{\alpha + \beta} f^{(\alpha)}(x) p^{(\beta)}(x).
	\end{equation}
	Whenever $\abs{\alpha + \beta} = u$ is odd, $I_{\alpha + \beta} = 0$; as a result $f_u = 0$ for all $u$ odd, and
	\begin{equation*}
	\Delta_rf(x) = \sum_{t = 1}^{\floor{(s - 1)/2}} f_{2t}(x)r^{2t} + O\bigl(Mp_{\max}r^s\bigr).
	\end{equation*}
	When $\abs{\alpha + \beta} = 2$, $I_{\alpha + \beta}$ is non-zero if and only if either $\alpha = \beta = e_i$ for some $i \in [d]$, or $\alpha = 2e_i$ and $\beta = 0$; here we write $e_i$ for the $i$th standard basis vector of $\Reals^d$. So the leading term becomes
	\begin{equation*}
	f_2(x) = -\sum_{i = 1}^{d} I_{2e_i}\Bigl\{\frac{1}{2}f^{(2e_i)}(x)p(x) + f^{(e_i)}(x)p^{(e_i)}(x)\Bigr\} = I_{2e_1} \cdot \Delta_pf(x).
	\end{equation*}
	Finally, if $\abs{\alpha + \beta} = u \geq 4$ and $\abs{\beta} \leq u - 1$, then $f^{(\alpha)} \in C^{s - \abs{\alpha}}(\Xset,M) \subset C^{s- u}(\Vset_r,M)$ and  $p^{(\beta)} \in C^{(s - 1) - \abs{\beta}}(\Xset,M) \subset C^{s - u}(\Vset_r,M)$. Along with~\eqref{eqn:expected_difference_holder_2}, this verifies $f_u \in C^{s - u}(\Vset_r,cM)$ for a constant $c$ which does not depend on $f$ or $M$, and completes the proof of~\eqref{eqn:expected_difference_holder_0} in the base case.

	Now for the induction step. Suppose $f \in C^s(\Vset,M)$ for $s \geq 2q + 3$, and that $B(x,r) \subset \Vset_{qr}$. Then by the inductive hypothesis
	\begin{align}
	\bigl(\Delta_r^{q + 1}f\bigr)(x) & = \bigl(\Delta_r \circ \Delta_r^qf)(x) \nonumber \\
	& = r^{2q} I_{2e_1}^q \bigl(\Delta_r \circ \Delta_p^q f\bigr)(x) + \sum_{t = q + 1}^{\floor{(s - 1)/2}} r^{2t} \bigl(\Delta_rf_{2t}\bigr)(x) + O\bigl(M p_{\max}r^s\bigr) \label{pf:expected_difference_holder_3}
	\end{align}
	where we have used the fact that $\Delta_r$ is a bounded---in the sense that $\norm{\Delta_r f}_{\Leb^{\infty}} \leq 2 \norm{f}_{\Leb^{\infty}}$---linear operator. In the first term on the right hand side of \eqref{pf:expected_difference_holder_3}, the iterated weighted Laplace operator $\Delta_p^qf \in C^{s - 2q}(\Xset,M) \subset C^3(\Xset,M)$, so that we may apply the inductive hypothesis and derive
	\begin{align}
	r^{2q} I_{2e_1}^q \bigl(\Delta_r \circ \Delta_p^q f\bigr)(x) & = r^{2(q + 1)} I_{2e_1}^{q + 1} \bigl(\Delta_p \circ \Delta_p^{q}f\bigr)(x) + \sum_{t = 2}^{\floor{(s - 2q - 1)/2}} r^{2(t + q)} I_{2e_1}^q f_{2t + 2q}(x) + r^{2q}O\bigl(Mp_{\max}r^{s - 2q}\bigr) \nonumber \\ 
	& = r^{2(q + 1)} I_{2e_1}^{q + 1} \bigl(\Delta_p^{q + 1}f\bigr)(x) + \sum_{h = 2(q + 1)}^{\floor{(s - 1)/2}} r^{2h} I_{2e_1}^q f_{2h}(x) + O\bigl(Mp_{\max}r^{s}\bigr) \label{pf:expected_difference_holder_4}
	\end{align}
	where the functions $f_u$ change between the first and second lines. For each term inside the sum on the right hand side of~\eqref{pf:expected_difference_holder_3}, since $f_{2t} \in C^{s - 2t}(\Vset_{qr},M)$ and $B(x,r) \subseteq \Vset_{qr}$, the inductive hypothesis implies
	\begin{align}
	r^{2t} \bigl(\Delta_rf_{2t}\bigr)(x) & = r^{2t}\biggl\{ r^2 I_{2e_1} \bigl(\Delta_pf_{2t}\bigr)(x) + \sum_{h = 2}^{\floor{(s - 2t - 1)/2}} r^{2h}f_{2h + 2t}(x) + O\bigl(M p_{\max} r^{s - 2t}\bigr) \biggr\} \nonumber \\
	& = r^{2(t + 1)} I_{2e_1} \bigl(\Delta_pf_{2t}\bigr)(x) + \sum_{h = t + 2}^{\floor{(s - 1)/2}} r^{2h}f_{2h}(x) + O\bigl(M p_{\max} r^{s}\bigr) \nonumber \\
	& = \sum_{h = t + 1}^{\floor{(s - 1)/2}} r^{2h}f_{2h}(x) + O\bigl(M p_{\max} r^{s}\bigr),\label{pf:expected_difference_holder_5}
	\end{align}
	where the last equality follows by setting $f_{2{t + 1}} = I_{2e_1}\Delta_pf_{2t}$. 
	Combining~\eqref{pf:expected_difference_holder_3}-\eqref{pf:expected_difference_holder_5}, we have that~\eqref{eqn:expected_difference_holder_0} holds for the $(q + 1)$st expected difference operator, completing the inductive step and thus the proof of the Lemma.
\end{proof}

\end{document}

### 2020-09-04 notes/graph_regression2.tex
Many changes to the graph regression paper. Most fundamentally, I have added a section in which we go through our results on graph Sobolev classes: we will see if this was a good choice.

@@
 \section{Introduction}
+\label{sec:introduction}
 
@@
 where $\kappa \in \{1,...,n\}$ is a tuning parameter.
-(Here $\dotp{u}{v}_n = \frac{1}{n}\sum_{i = 1}^{n} u_i v_i$ for vectors $u,v \in \Reals^n$.) The Laplacian smoothing estimator $\wt{f}_{\LS}$ \citep{smola2003} is given by
+(Here $\dotp{u}{v}_n = \frac{1}{n}\sum_{i = 1}^{n} u_i v_i$ for vectors $u,v \in \Reals^n$.) The Laplacian smoothing estimator $\wt{f}_{\LS}$ \citep{smola2003} is a penalized least square estimator, given by
 \begin{equation}
@@
 \end{equation}
-where $\rho > 0$ acts a tuning parameter. Then, assuming~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are reasonable estimators of $f_0$, the size of the statistics
+where $\rho > 0$ acts a tuning parameter $f^T \Lap_{{n,r}}^s f$.
+Then, assuming~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are reasonable estimators of $f_0$, the squared empirical norms
 \begin{equation}
@@
 \end{equation}
-can be used to test whether or not $f_0 = 0$.
+are in turn reasonable statistics to test whether or not $f_0 = 0$.
 
@@
 		\item Computational efficiency and stability.
-		\item Domain and density adaptivity. Easily generalizable. 
+		\item Domain and density adaptivity.
+		\item Easily generalizable.
 		\item Practical success.
@@
 
+\section{Graph Sobolev classes}
+\label{sec:graph_sobolev_classes}
+
+\begin{itemize}
+	\item For $s,M > 0$, define the order-$s$ graph Sobolev ball to be
+	\begin{equation}
+	\label{eqn:graph_sobolev_ball}
+	H^s(G_{n,r},M) = \set{f \in \Reals^n: f^T \Lap_{n,r}^s f \leq M}
+	\end{equation}
+	where we recall that the \emph{graph Sobolev seminorm} $f^T \Lap_{{n,r}}^s f$ is precisely the penalty term in~\eqref{eqn:laplacian_smoothing}. 
+	To understand the relevance of the graph Sobolev ball to the analysis of our graph-based procedures, let us reason again by analogy with the continuum world.
+The statistical analysis of continuum-time methods such as Fourier series projection and smoothing splines rests on two facts about the continuum classes $\mc{H}^s(\Xset)$: first, the \textit{a priori} assumption that the regression function $f_0$ belongs to an appropriate Sobolev ball $\mc{H}(\Xset,M)$, and second that the asymptotic decay of the eigenvalues of the continuum Laplace operator $\Delta$ obey Weyl's law, meaning $\lambda_k(\Delta) \asymp \abs{k}^{2s}$ for $k \in \mathbb{Z}^d$.
+In this section, we will show that both these statements have appropriate analogues in the graph setting.
+Roughly speaking, we will show first that for functions $f$ belonging to $\mc{H}^s(\Xset, M)$, the evaluations $f = (f(X_1),\ldots,f(X_n)) \in \Reals^n$ satisfy the upper bound
+	\begin{equation}
+	\label{eqn:graph_sobolev_seminorm_asymptotic}
+	f^T \Lap_{n,r}^s f \lesssim M(n^{s + 1}r^{s(d + 2)})
+	\end{equation}
+	and second that the eigenvalues $\lambda_k(G_{n,r})$ satisfy the lower bound
+	\begin{equation*}
+	\bigl[\lambda_k(G_{n,r})\bigr]^s = \lambda_k(\Lap_{{n,r}}^s) \gtrsim k^{2s/d} (n^s r^{s(d + 2)});
+	\end{equation*}
+	both statements will hold with high probability.
+These results elucidate why our graph-based procedures ``work'' as well as their continuum counterparts---in the sense of being statistically rate-optimal over (continuum) Sobolev classes.
+They also clarify when our analysis does and does not apply---in terms of various degrees of smoothness $s$ and dimensions $d$.
+\end{itemize}
+
+\subsection{Graph Sobolev seminorm}
+We will state and prove three different results involving the graph Sobolev semi-norm $f_0 \Lap_{n,r}^s f_0$.
+In each case, we will derive that the asymptotic relation~\eqref{eqn:graph_sobolev_seminorm_asymptotic} holds with high probability, whenever $r \to 0$ sufficiently slowly as $n \to \infty$.
+The difference between the three results will be the assumptions we make, and the resulting rate of permissible decay for the radius $r$.
+
+\subsubsection{First-order graph Sobolev semi-norm.}
+We begin by upper bounding $f_0^T \Lap_{n,r} f_0$ under the assumption $f_0 \in H^1(\Xset,M)$.
+This will be the easiest of the three Lemmas to prove, while still giving some sense of how the proofs develop in the more complicated cases.
+The upper bound will depend on the kernel $K$; we assume throughout that $K:[0,\infty) \to [0,\infty)$ is a $K_{\max}$-Lipschitz function supported on $[0,1]$, and without loss of generality that
+\begin{equation*}
+\int_{\Reals^d} K(\norm{z}) \,dz = 1.
+\end{equation*}
+\begin{lemma}
+	\label{lem:first_order_graph_sobolev_seminorm}
+	Suppose $\Xset$ is a Lipschitz domain, and that $p(x) \leq p_{\max}$ for all $x \in \Xset$.
+Then, for any $f \in H^1(\Xset,M)$, there exists a constant $C > 0$ which may depend only on $\Xset$ and $d$ such that
+	\begin{equation}
+	f^T \Lap_{n,r} f \leq \frac{C}{\delta} n^2 r^{d + 2} M^2
+	\end{equation}
+	with probability at least $1 - \delta$.
+\end{lemma}
+
+\subsubsection{General graph Sobolev semi-norm}
+
+When $s > 1$, in order to obtain the desired upper bound on the graph Sobolev semi-norm we must make some additional assumptions regarding the behavior of $f_0$ near the boundary of $\Xset$, as well as on the smoothness of the density function $p$.
+\begin{theorem}
+	\label{thm:graph_sobolev_seminorm}
+	For some integer $s > 1$, suppose that the density $p \in C^{s - 1}(\Xset,p_{\max})$ and $f \in H_0^s(\Xset,M)$, and additionally that the domain $\Xset$ is Lipschitz.
+Let $r = r(n)$ be a sequence of connectivity radii satisfying $r(n) \to 0$ and $r(n) \gtrsim n^{-1/(2(s - 1) +d)}$ as $n \to \infty$.
+Then for all $n$ sufficiently large, there exists a constant $C > 0$ which may depend only on $\Xset$ and $d$ such that
+	\begin{equation}
+	\label{eqn:graph_sobolev_seminorm}
+	f^T \Lap_{n,r}^s f \leq \frac{C}{\delta} n^{s + 1} r^{s(d + 2)}M^2
+	\end{equation}
+	with probability at least $1 - \delta$.
+\end{theorem}
+It is worth pausing to make a few remarks.
+\begin{itemize}
+	\item \textbf{Bounded norm vs. bounded semi-norm.}
+	\item \textbf{Markov's inequality, as opposed to exponential concentration.}
+	\item \textbf{Simpler proof and stronger concentration under Holder assumption.}
+	\item \textbf{Explain why~\eqref{eqn:graph_sobolev_seminorm} should be familiar, and a lower bound on $r$ expected.} The scaling (in $n$, $r$, and $M$) on the right hand side of~\eqref{eqn:graph_sobolev_seminorm} will be familiar to the informed reader.
+For a sufficiently regular function $f \in C^3(\Xset,M)$ and density $p \in C^1(\Xset,p_{\max})$, it is known \textcolor{red}{(references)} that $\frac{1}{nr^{d + 2}}\Lap_{n,r}f \to \Delta_pf$ for \textcolor{red}{$r \geq ???$}.
+\textcolor{red}{A recursive argument shows that when $f \in C^{2s + 1}(\Xset)$ and $p \in C^{2s}(\Xset)$ are sufficiently regular and $r$ is sufficiently large},
+	\begin{equation*}
+	\Bigl(\frac{1}{nr^{d + 2}}\Lap_{n,r}\Bigr)^s f \to \Delta_p^sf \Longrightarrow \frac{1}{n^{s + 1} r^{s(d + 2)}} f^T \Lap_{n,r}^s f \to \int_{\Xset} f(x) \cdot  \Delta_p^sf(x) p(x) \,dx \leq c M^2.
+	\end{equation*}
+	\item \textbf{Explain why the actual proof is more subtle.} In fact, the actual proof of Theorem~\ref{thm:graph_sobolev_seminorm} is rather more involved.
+Among other subtleties we highlight the mismatch, in the previous remark, between the number of smooth derivatives $f$ possesses ($2s + 1$) and the degree of the Laplacian operator $\Lap_{{n,r}}^s$ $(s)$; this is unsuitable for our purposes.
+A more careful analysis implies~\eqref{thm:graph_sobolev_seminorm}, but note well that~\eqref{eqn:graph_sobolev_seminorm} is only an upper bound rather than a statement about consistency.
+This is the price we pay for insisting that the assumed number of smooth derivatives match the degree of the Laplacian; fortunately an upper bound will be sufficient for our purposes.
+\end{itemize}
+
+\subsubsection{Graph Sobolev semi-norm under manifold assumption.}
+
+\textcolor{red}{(TODO)}
+
+\subsection{Graph Laplacian eigenvalues}
+
+\textcolor{red}{(TODO)}
+
 \section{Minimax optimal graph Laplacian methods}
@@
 In this section, we will establish that the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and the resulting test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test}, achieve the standard minimax optimal estimation and hypothesis testing rates, respectively.
+Pleasingly, these results will go through with essentially no additional regularity conditions---additional, that is, to~\ref{asmp:bounded_density}---on either the regression function $f_0$ or the density $p$.
+The story will also be complete: we will obtain the ``right'' rates for all regimes--that is, all values of $s$ and $d$---for which we would expect to get these rates.
 
@@
 The underlying phenomenon is similar to that elucidated by the seminal papers~\cite{birge1993} and~\cite{birge1998}: when the dimension $d$ is large enough, the parameter space is too large, and (penalized) least squares approaches are no longer minimax optimal.
-It is interesting to note that we see a gap emerge between the graph- (Laplacian smoothing) and continuum (tensor product spline) estimator when $d = 3$: the former continues to be minimax optimal, whereas the latter no longer is.
+It is interesting to note that we see a gap emerge between the graph- (Laplacian smoothing) and continuum (smoothing spline) estimator when $d = 3$: the former continues to be minimax optimal, whereas the latter is not even well-defined.
 We should also mention that in a related context (estimation over a $d$-dimensional lattice graph), \cite{sadhanala2016} require the same restriction $d \leq 4$.

### 2020-09-04 notes/weekly_notes/8.26.20.NotesForWeek.tex
Many changes to the graph regression paper. Most fundamentally, I have added a section in which we go through our results on graph Sobolev classes: we will see if this was a good choice.

@@
 \section{Bounds on the graph Sobolev semi-norm}
-In this section we will state and prove three different Lemmas involving the graph Sobolev semi-norm $f_0 \Lap_{n,r}^s f_0$. In each Lemma, we will derive that the asymptotic relation
+In this section we will state and prove three different results involving the graph Sobolev semi-norm $f_0 \Lap_{n,r}^s f_0$. In each Lemma, we will derive that the asymptotic relation
 \begin{equation*}
@@
 holds with high probability, whenever $r \to 0$ sufficiently slowly as $n \to \infty$.
-The difference between the three Lemmas will be the assumptions we make, and the resulting permissible range for $r$.
-Throughout, we will make heavy use of Taylor's Theorem, which we briefly recall.
-For an open set $U \subseteq \Reals^d$ and a function $g \in C^{\infty}(U)$, it holds that for all $x,x' \in U$, and for every $s \geq 1$,
+The difference between the three results will be the assumptions we make, and the resulting range of permissible values for the radius $r$.
+
+\paragraph{Some facts from calculus.}
+
+The analysis in this section will frequently appeal to Taylor's Theorem, which we now briefly recall in a form convenient for our purposes.
+For an open set $U \subseteq \Reals^d$ and a function $g \in C^{s}(U)$, it holds that for all $x,x' \in U$,
 \begin{equation*}
-g(x') = \sum_{\abs{\alpha} = 0}^{s - 1} g^{(\alpha)}(x) \cdot (x' - x)^{\alpha}  + R_g^{(s)}(x',x)
+g(x') = \sum_{\abs{\alpha} = 0}^{s - 1} g^{(\alpha)}(x) \cdot (x' - x)^{\alpha}  + \sum_{\abs{\alpha} = s}R_g^{(\alpha)}(x,x') \cdot (x' - x)^{\alpha}
 \end{equation*}
-where for concision we have written 
+where for $\alpha \in \mathbb{N}^d$ we use the multiindex notation $x^{\alpha} = \prod_{i = 1}^{d} x_i^{\alpha_i}$, and
 \begin{equation*}
-g^{(\alpha)}(x) = \frac{1}{\abs{\alpha}!} \bigl(D^{\alpha}g\bigr)(x),~~ R_g^{(s)}(x',x) = \frac{1}{s!} \sum_{\abs{\alpha} = s} (x' - x)^{\alpha} \int_{0}^{1} (1 - t)^{\abs{\alpha}}  g^{(\alpha)}\bigl(x + t(x' - x)\bigr) \,dt
+g^{(\alpha)}(x) = \frac{1}{\abs{\alpha}!} \bigl(D^{\alpha}g\bigr)(x),~~ R_g^{(\alpha)}(x,x') = \frac{1}{\abs{\alpha}!} \int_{0}^{1} (1 - t)^{\abs{\alpha}}  g^{(\alpha)}\bigl(x + t(x' - x)\bigr) \,dt,
 \end{equation*}
-and $x^{\alpha} = \prod_{i = 1}^{d} x_i^{\alpha_i}$ for $x,\alpha \in \Reals^d$.
+where we note that the integral form of the remainder is well-defined by Rademacher's theorem.
+To keep our equations as compact as possible, we also introduce the following notation.
+For a remainder term $R: U \times U \to \Reals$, let
+\begin{align*}
+\Bigl(\Ebb_{\alpha}[R]\Bigr)(x) & := \int_{U} (x' - x)^{\alpha} R(x,x') K_r(x,x') p(x') \,dx',~~ && \Ebb_{\alpha}(x) := \Bigl(\Ebb_{\alpha}[1]\Bigr)(x) \\
+\Bigl(\Ibb_{\alpha}[R]\Bigr)(x) & := \int_{U} (x' - x)^{\alpha} R(x,x') K_r(x,x') \,dx',~~ && \Ibb_{\alpha}(x)  := \Bigl(\Ibb_{\alpha}[1]\Bigr)(x);
+\end{align*}
+and
+\begin{equation*}
+I_{\alpha} := \int_{B(0,1)} (z)^{\alpha} K\bigl(\norm{z}\bigr) \,dz
+\end{equation*}
+where $B(0,1) \subseteq \Reals^d$ is the unit ball centered at the origin. (It will always be clear from context what the domain $U$ of integration is, and so we suppress it in our notation.)
+
+Let $V \subset U$ satisfy $B(x,r) \subset U$ for all $x \in V$.
+Before moving on, we will record four facts that will later be of use to us:  first, that $\Ibb_{\alpha}(x) = r^{\abs{\alpha} + d} I_{\alpha}$; second that $I_{\alpha} = 0$ unless $\alpha_i$ is even for all $i \in [d]$; third, that
+\begin{equation}
+\label{eqn:integrated_remainder_term_sobolev}
+\Bigl\|\Ebb_{\alpha}\bigl[R_g^{(\alpha)}\bigr]\Bigr\|_{\Leb^2(V)} \leq p_{\max} K_{\max} r^{\abs{\alpha} + d} \abs{g}_{H^{\abs{\alpha}}(U)}
+\end{equation}
+and fourth that
+\begin{equation}
+\label{eqn:integrated_remainder_term_holder}
+\Bigl\|\Ebb_{\alpha + \beta}\bigl[R_g^{(\beta)}\bigr]\Bigr\|_{\Leb^{\infty}(V)} \leq p_{\max} K_{\max} r^{\abs{\alpha + \beta} + d} |g|_{C^{\abs{\beta}}(U)}
+\end{equation}
+The proofs of~\eqref{eqn:integrated_remainder_term_sobolev} and~\eqref{eqn:integrated_remainder_term_holder} are straightforward; for purposes of completeness we provide them in Section~\ref{subsec:proof_integrated_remainder_terms}.
+Finally, we note that for $g \in C^{s}(U)$ and $f \in H^s(U)$, 
+\begin{equation}
+\label{eqn:product_rule_sobolev}
+\bigl\|f \cdot g\bigr\|_{H^s(U)} \leq C \norm{g}_{C^s(U)} \norm{f}_{H^s(U)}
+\end{equation}
+for a constant $C$ which depends only on $s$, but which we will not bother to identify further.
 
@@
 This will be the easiest of the three Lemmas to prove, while still giving some sense of how the proofs develop in the more complicated cases.
-The upper bound will depend on the kernel $K$; we assume throughout that $K:[0,\infty) \to [0,\infty)$ has support on $[0,1]$ with
+The upper bound will depend on the kernel $K$; we assume throughout that $K:[0,\infty) \to [0,\infty)$ is a $K_{\max}$-Lipschitz function supported on $[0,1]$, and without loss of generality that
 \begin{equation*}
-\int_{\Reals^d} K(\norm{z}) \,dz = 1,
+\int_{\Reals^d} K(\norm{z}) \,dz = 1.
 \end{equation*}
-and we let
-\begin{equation*}
-I_{\alpha} := \int_{B(0,1)} (z)^{\alpha} K\bigl(\norm{z}\bigr) \,dz
-\end{equation*}
-where $B(0,1) \subseteq \Reals^d$ is the unit ball centered at the origin.
 \begin{lemma}
@@
 
-\subsection{General graph Sobolev semi-norm}
+\subsection{Order-$s$ graph Sobolev semi-norm}
 
@@
 	\label{thm:graph_sobolev_seminorm}
-	Suppose $\Xset $ is a Lipschitz domain, and that the density $p \in C^{s - 1}(\Xset,p_{\max})$.
-Then, for any $f \in H_0^s(\Xset,M)$ there exists a constant $C > 0$ which may depend only on $X$ and $d$ such that
+	For some integer $s > 1$, suppose that the density $p \in C^{s - 1}(\Xset,p_{\max})$ and $f \in H_0^s(\Xset,M)$, and additionally that the domain $\Xset$ is Lipschitz.
+Let $r = r(n)$ be a sequence of connectivity radii satisfying $r(n) \to 0$ and $r(n) \gtrsim n^{-1/(2(s - 1) +d)}$ as $n \to \infty$.
+Then for all $n$ sufficiently large, there exists a constant $C > 0$ which may depend only on $\Xset$ and $d$ such that
 	\begin{equation}
@@
 \end{theorem}
-The proof of Lemma~\ref{thm:graph_sobolev_seminorm} will be considerably longer than that of Lemma~\ref{lem:first_order_graph_sobolev_seminorm}.
+A few remarks:
+\begin{itemize}
+	\item \textbf{Explain why~\eqref{eqn:graph_sobolev_seminorm} should be familiar, and a lower bound on $r$ expected.} The scaling (in $n$, $r$, and $M$) on the right hand side of~\eqref{eqn:graph_sobolev_seminorm} will be familiar to the informed reader.
+For sufficiently regular functions $f$, it is known \textcolor{red}{(references)} that $\frac{1}{nr^{d + 2}}\Lap_{n,r}f \to \Delta_pf$ for \textcolor{red}{$r \geq ???$}.
+In that light, it is perhaps reasonable to hope that when $r$ is sufficiently large, some recursive argument yields
+	\begin{equation*}
+	\Bigl(\frac{1}{nr^{d + 2}}\Lap_{n,r}\Bigr)^s f \to \Delta_p^qf \Longrightarrow \frac{1}{n^{s + 1} r^{s(d + 2)}} f^T \Lap_{n,r}^s f \to \int_{\Xset} f(x) \cdot  \Delta_p^qf(x) p(x) \,dx \leq c M^2.
+	\end{equation*}
+	\item \textbf{Explain why the actual proof is more subtle.}
+\end{itemize}
+The proof of Theorem~\ref{thm:graph_sobolev_seminorm} will be considerably longer than that of Lemma~\ref{lem:first_order_graph_sobolev_seminorm}.
 In anticipation of this proof, it will be helpful to introduce some notation.
@@
 \end{equation*}
-and for $k \in [n]^q$ for some $q \in \mathbb{N}$, define the iterated difference operator $D_{jk}f: \Xset \to \Reals$ via the recursive formula
+and for $k \in [n]^q$ for some $q \in \mathbb{N}$, define the iterated difference operator $D_{jk}f: \Xset \to \Reals$ via the recursive relation
 \begin{equation*}
@@
 Consequently, a few standard computations show that the graph Sobolev semi-norm can be written as 
-\begin{equation*}
-f^T \Lap_{n,r}^s f = 
+\begin{equation}
+\label{eqn:graph_sobolev_seminorm_1}
+f^T \Lap_{n,r}^s f = r^{ds} \cdot
 \begin{dcases}
@@
 \end{dcases}
-\end{equation*}
+\end{equation}
 where $\bigl(d_jf\bigr)(x) := f(X_j) - f(x)$ is the simple difference operator.
 
-In order to prove~\eqref{eqn:graph_sobolev_seminorm}, we will proceed as in Lemma~\ref{lem:first_order_graph_sobolev_seminorm} by first upper bounding $\Ebb[f^T \Lap_{n,r}^s f]$ and then applying Markov's inequality.
-Obtaining the desired upper bound on $\Ebb[f^T \Lap_{n,r}^s f]$ is split into \textcolor{red}{N} Lemmas, Lemmas \textcolor{red}{A - B}.
-Lemma~\ref{lem:expected_difference_operator} controls the expectation of the ``leading terms'' $\Ebb[D_kf(x)]$, when $k \in (n)^q$ and $x$ is sufficiently in the interior of $\Xset$.
+In order to prove~\eqref{eqn:graph_sobolev_seminorm}, we will proceed as we did in proving Lemma~\ref{lem:first_order_graph_sobolev_seminorm}: we first upper bound $\Ebb[f^T \Lap_{n,r}^s f]$, and then apply Markov's inequality.
+From the representation~\eqref{eqn:graph_sobolev_seminorm_1} we gather that to analyze the expected graph Sobolev seminorm, it suffices to evaluate the expectation of products of iterated difference operators, and we do the latter in Lemmas \textcolor{red}{A - B} which we now state.
+
+Let $\Delta_r$ be the expected weighted difference operator
+\begin{equation}
+\Delta_rf(x) := \Ebb\Bigl[\bigl(D_if\bigr)(x)\Bigr] = -\frac{1}{r^d}\int_{\Xset} \bigl(f(x') - f(x)\bigr) K_r(x',x) p(x') \,dx
+\end{equation}
+and define $\Delta_r^q = \Delta_r \circ \cdots \circ \Delta_r$, with the operator $\Delta_r$ being composed a total of $q$ times.
+Lemma~\ref{lem:expected_difference_sobolev} shows that when $k \in (n)^q$, the iterated expected weighted difference operator $\Delta_r^qf$ approximates the iterated Laplace operator $\Delta_p^qf$ (after an appropriate rescaling), averaged over points $x$ sufficiently far from the boundary of $\Xset$.
+Lemma~\ref{lem:expected_difference_boundary} shows that when $k \in (n)^q$, the zero-trace condition implies that $\Delta_r^qf$ is itself small, averaged over points $x$ near the boundary of $\Xset$.
+In Lemmas~\ref{lem:expected_difference_lower_order_1} and~\ref{lem:expected_difference_lower_order_2}, we take care of those remaining terms, where there is at least one repeated index in the iterated weighted difference operator.
+
 \begin{lemma}
-	\label{lem:expected_difference_operator}
+	\label{lem:expected_difference_sobolev}
+	Let $q \geq 1$ and $s \geq 2q + 1$ be integers.
+The iterated expected weighted difference operator  $\Delta_r^qf(x) = \Ebb[D_kf(x)]$ for any $k \in (n)^q$.
+If $f \in H^s(\Xset,M)$ and $p \in C^{s-1}(\Xset,M)$,
+	\begin{equation*}
+	\Bigl\|\Delta_r^qf - \bigl(r^{2q} I_{2e_1}^{q}\bigr) \Delta_{p}^qf\Bigr\|_{\Leb^2(\Xset_{qr})} \leq CM \cdot
+	\begin{dcases}
+	r^{2q + 1},& ~~\textrm{if $s = 2q + 1$} \\
+	r^{2q + 2},& ~~\textrm{if $s = 2q + 2$}
+	\end{dcases}
+	\end{equation*}
+	where $\Xset_{qr} = \{x \in \Xset: B(x,qr) \subset \Xset\}$, and we recall that $\Delta_pf = -\frac{1}{2p}\mathrm{div}(p^2 \nabla f)$ is the density weighted Laplace operator.
 \end{lemma}
+\begin{lemma}
+	\label{lem:expected_difference_boundary}
+	Let $q \geq 1$ and $s \geq 1$ be integers. If $f \in H_0^s(\Xset,M)$ and \textcolor{red}{$p$ satisfies (P2)}, then for all $r > 0$ sufficiently small,
+	\begin{equation*}
+	\norm{\Delta_r^qf}_{\Leb^2(\Xset^{qr})} \leq cMr^s
+	\end{equation*}
+	where $\Xset^{qr} = \Xset \setminus \Xset_{qr}$.
+\end{lemma}
+\begin{lemma}
+	\label{lem:expected_difference_lower_order_1}
+	Let $I = \abs{k \cup \ell \cup i} < 2q + 1$ for given index vectors $k,\ell \in [n]^q$ and index $i \in [n]$. Then for any $f \in H^1(\Xset)$,
+	\begin{equation*}
+	\Bigl|\Ebb\Bigl[\bigl(D_kf\bigr)(X) \cdot \bigl(D_{\ell}f\bigr)(X)\Bigr]\Bigr| \leq C r^{2 + d(I - (s + 1))} \abs{f}_{H^1(\Xset)}^2
+	\end{equation*}
+	for a constant $C$ which does not depend on $f$.
+\end{lemma}
+A few remarks:
+\begin{itemize}
+	\item In~\textcolor{red}{Garcia-Trillos19}, results similar to Lemma~\ref{lem:expected_difference_sobolev} were shown, as part of the proof of pointwise convergence of the graph Laplacian $\Lap_{n,r}$ to $\Delta_p$.
+Those results specialize to the case $q = 1$ and $s = 3$, and assume the function $f$ is bounded in the relevant Holder rather than Sobolev norm, sufficient for their purposes but not for our own.
+	\item When the function $f$ lacks sufficiently regularity---i.e.
+when it has two or fewer bounded derivatives---we can no longer obtain an estimate of  $\Delta_rf$ in terms of $\Delta_pf$.
+However, we can still show that $\Delta_rf$ is itself bounded.
+For an open set $V \subseteq \Xset$, if $f \in H^{s}(V,M)$---and when $s = 2$, $p \in C^{1}(\Xset,M)$---then
+	\begin{equation}
+	\label{eqn:norm_expected_difference_operator}
+	\norm{\Delta_rf}_{\Leb^2(V_r)} \leq c r^s \norm{f}_{H^s(V)}
+	\end{equation}
+	for $V_r = \set{x \in V: B(x,r) \subset V}$. We establish this fact, and bound a few related integrals, in Subsection~\ref{subsec:proof_boundedness_expected_difference_operator}.
+	\item \textbf{Discuss how the analysis in \textcolor{red}{Belkin12} could conceivably lead to similar conclusions as Lemma~\ref{lem:expected_difference_boundary} under a milder assumption than zero-trace.
+Explain why you didn't bother to go down this road.}
+\end{itemize}
+
+\begin{proof}[\textbf{Proof of Lemma~\ref{lem:expected_difference_sobolev}}.]
+	Without loss of generality we will assume $f \in C^{\infty}(\Xset)$ and $p \in C^{\infty}(\Xset)$, as justified in the proof of Theorem~\ref{thm:graph_sobolev_seminorm}.
+
+	We will actually prove a more general claim.
+Let $f \in C^{\infty}(\Xset)$ additionally satisfy $f \in H^s(\Vset,M)$ for an open set $\Vset \subset \Xset$, and let $s \geq 2q + 1$.
+Letting $\Vset_{qr} = \{x \in \Vset: B(x,qr) \subseteq \Vset\}$, we will show that there exist functions $f_{2t} \in H^{s - 2t}(\Vset_{qr},CM)$ for $t = q + 1,\ldots,\floor{(s - 1)/2}$ such that
+	\begin{equation}
+	\label{pf:expected_difference_sobolev_0}
+	\Bigl\| \Delta_r^q f  - \bigl(r^{2q} I_{2e_1}^q\bigr) \Delta_p^q f - \sum_{t = q + 1}^{\floor{(s - 1)/2}} r^{2t}f_{2t} \Big\|_{\Leb^2(\Vset_{qr})} \leq C M r^s
+	\end{equation}
+	for a constant $C$ which depends only on $\Vset$, $d$, $s$, and $p_{\max}$.
+We note that in our proof the functions $f_u$ may differ from line to line, but they will always satisfy $f_u \in H^{s - u}(\Vset_{qr},M)$.
+To see that~\eqref{pf:expected_difference_sobolev_0} implies the claim of the Lemma, simply take $\Vset = \Xset$, and note that when $s = 2q + 1$ or $s = 2q + 2$, the sum on the left hand side of~\eqref{pf:expected_difference_sobolev_0} is empty.
+
+	We proceed by induction on $q$. When $q = 1$, the expected difference operator
+	\begin{equation*}
+	\Delta_r^qf(x) = \Delta_rf(x) = -\frac{1}{r^d} \int_{\Xset} \bigl(f(x') - f(x)\bigr) K_r(x',x) p(x') \,dx' =  -\frac{1}{r^d} \int_{\Vset} \bigl(f(x') - f(x)\bigr) K_r(x',x) p(x') \,dx'
+	\end{equation*}
+	where we may restrict the integral to $\Vset$ since $B(x,r) \subset \Vset$ and $K$ has support on $[0,1]$.
+Taking Taylor expansions of $f$ and $p$ around $x' = x$, and writing things in our compact notation, we have
+	\begin{align}
+	\bigl(\Delta_rf\bigr)(x) & = -\frac{1}{r^d} \Biggl\{ \sum_{\abs{\alpha} = 1}^{s - 1} f^{(\alpha)}(x) \Ebb_{\alpha}(x) + \sum_{\abs{\alpha} = s}\Bigl(\Ebb_{\alpha}\bigl[R_f^{(\alpha)}\bigr]\Bigr)(x) \Biggr\}  \nonumber\\
+	& = -\frac{1}{r^d} \Biggl\{ \sum_{\abs{\alpha} = 1}^{s - 1} f^{(\alpha)}(x) \biggl[ \sum_{\abs{\beta} = 0}^{s - 2} p^{(\beta)}(x) \Ibb_{\alpha + \beta}(x) + \sum_{\abs{\beta} = s - 1} \Bigl(\Ibb_{\alpha + \beta}\bigl[R_p^{(\beta)}\bigr]\Bigr)(x)  \biggr] + \sum_{\abs{\alpha} = s}\Bigl(\Ebb_{\alpha}\bigl[R_f^{(\alpha)}\bigr]\Bigr)(x) \Biggr\} \nonumber \\
+	& = - \sum_{\abs{\alpha} = 1}^{s - 1}  \sum_{\abs{\beta} = 0}^{s - 2}  f^{(\alpha)}(x) p^{(\beta)}(x) I_{\alpha + \beta} r^{\abs{\alpha + \beta}} - \frac{1}{r^d} \Biggl\{\sum_{\abs{\alpha} = 1}^{s - 1} \sum_{\abs{\beta} = s - 1} f^{(\alpha)}(x)\Bigl(\Ibb_{\alpha + \beta}\bigl[R_p^{(\beta)}\bigr]\Bigr)(x)+ \sum_{\abs{\alpha} = s}\Bigl(\Ebb_{\alpha}\bigl[R_f^{(\alpha)}\bigr]\Bigr)(x) \Biggr\} \nonumber \\
+	& = \sum_{u = 1}^{2s - 3} f_u(x) r^{u} - \frac{1}{r^d} \Biggl\{\sum_{\abs{\alpha} = 1}^{s - 1} \sum_{\abs{\beta} = s - 1} f^{(\alpha)}(x)\Bigl(\Ibb_{\alpha + \beta}\bigl[R_p^{(\beta)}\bigr]\Bigr)(x)+ \sum_{\abs{\alpha} = s}\Bigl(\Ebb_{\alpha}\bigl[R_f^{(\alpha)}\bigr]\Bigr)(x) \Biggr\} \label{pf:expected_difference_sobolev_1}
+	\end{align}
+	where the final equality follows by putting
+	\begin{equation}
+	\label{pf:expected_difference_sobolev}
+	f_u(x) = - \sum_{\abs{\alpha} + \abs{\beta} = u} I_{\alpha + \beta} f^{(\alpha)}(x) p^{(\beta)}(x),
+	\end{equation}
+    where the sum is over indices $1 \leq \abs{\alpha} \leq s - 1$ and $0 \leq \abs{\beta} \leq s - 2$.
+
+	The remainder terms---those inside the braces in~\eqref{pf:expected_difference_sobolev_1}--have $\Leb^2(V_r)$ norm on the order of at most $r^s$; when $\alpha = \abs{s}$
+	\begin{equation*}
+	\frac{1}{r^d} \Bigl\|\Ebb_{\alpha}\bigl[R_f^{(\alpha)}\bigr]\Bigr\|_{\Leb^2(\Vset_r)} \leq p_{\max} K_{\max} M r^s
+	\end{equation*}
+	follows directly from~\eqref{eqn:integrated_remainder_term_sobolev}, and when $\abs{\alpha} \geq 1$ and $\abs{\beta} = s - 1$,
+	\begin{equation*}
+	\frac{1}{r^d} \Bigl\|f^{(\alpha)} \cdot \Ibb_{\alpha + \beta}\bigl[R_p^{(\beta)}\bigr]\Bigr\|_{\Leb^2(\Vset_r)} \leq \frac{1}{r^d}\Big\|f^{(\alpha)} \Bigr\|_{\Leb^2(\Vset_r)} \Big\|\Ibb_{\alpha + \beta}\bigl[R_p^{(\beta)}\bigr]\Bigr\|_{\Leb^{\infty}(\Vset_r)} \leq p_{\max} K_{\max} M r^s
+	\end{equation*}
+	follows from~\eqref{eqn:integrated_remainder_term_holder}.
+
+	It remains to deal with the functions $f_u$.
+Whenever $\abs{\alpha + \beta} = u$ is odd, $I_{\alpha + \beta} = 0$; as a result $f_u = 0$ for all $u$ odd.
+When $u = 2$, $I_{\alpha + \beta}$ is non-zero only if either $\alpha = \beta = e_i$ for some $i \in [d]$, or $\alpha = 2e_i$ and $\beta = 0$.
+So the leading term becomes
+	\begin{equation*}
+	f_2(x) = \sum_{i = 1}^{d} I_{2e_i}\Bigl\{\frac{1}{2}f^{(2e_i)}(x)p(x) + f^{(e_i)}(x)p^{(e_i)}(x)\Bigr\} = I_{2e_1} \cdot \Delta_pf(x).
+	\end{equation*}
+	Next, if $4 \leq \abs{\alpha + \beta} = u \leq s$ and $\abs{\beta} \leq u - 1$, then 
+	\begin{equation*}
+	f^{(\alpha)} \in H^{s - \abs{\alpha}}(\Xset,M) \subseteq H^{s - u}(\Vset_r,M) ~~\textrm{and}~~ p^{(\beta)} \in C^{(s - 1) - \abs{\beta}}(\Xset,p_{\max}) \subset C^{s - u}(\Vset_r,p_{\max})
+	\end{equation*}
+	which along with~\eqref{pf:expected_difference_sobolev} verifies $f_u \in H^{s - u}(\Vset_r,cM)$ for a constant $c$ which does not depend on $f$ or $M$.
+Finally, if $\abs{\alpha + \beta} = u > s$ then
+	\begin{equation*}
+	f^{(\alpha)} \in \Leb^2(\Xset,M) \subseteq \Leb^2(\Vset_r,M) ~~\textrm{and}~~ p^{(\beta)} \in \Leb^{\infty}(\Xset,p_{\max}) \subseteq \Leb^{\infty}(\Vset_r,p_{\max})
+	\end{equation*}
+	and so $f_u \in \Leb^2(\Vset_r,cM)$. Hence we have proved~\eqref{pf:expected_difference_sobolev_0} in the base case $q = 1$.
+
+	Now for the induction step. Suppose $f \in H^s(\Vset,cM)$ for $s \geq 2(q + 1) + 1$. Since $\Delta_r^{(q + 1)}f = \Delta_r \circ \Delta_r^qf$, applying the inductive hypothesis to $\Delta_r^qf$ gives
+	\begin{equation}
+	\label{pf:expected_difference_sobolev_2}
+	\Bigl\| \Delta_r^{(q + 1)}f - r^{2q}I_{2e_1}^q\Delta_r(\Delta_p^qf) + \sum_{t = q + 1}^{\floor{(s - 1)/2}}r^{2t} \Delta_r(f_{2t}) \Bigr\|_{\Leb^2(\Vset_{(q + 1)r})} \leq c M r^s
+	\end{equation}
+	Looking at those terms $\Delta_r(f_{2t})$, when $2t \leq s  - 3$ then $f_{2t} \in H^{s - 2t}(\Vset_{qr},cM) \subset H^3(\Vset_{qr},cM)$ and it follows from the inductive hypothesis that
+	\begin{equation*}
+	\biggl\| r^{2t}\Delta_r(f_{2t}) - r^{2(t + 1)}I_{2e_1}\Delta_p(f_{2t}) - \sum_{h = 2}^{\floor{(s - 2t - 1)/2}}r^{2(t + h)}f_{2t + 2h}\biggr\|_{\Leb^2(\Vset_{(q + 1)r})} \leq c M r^{2t + (s  - 2t)}.
+	\end{equation*}
+	We note that $\Delta_p(f_{2t}) \in H^{s - 2(t + 1)}(\Vset_{qr},cM)$, and we may therefore write the above expression more succinctly,
+	\begin{equation}
+	\label{pf:expected_difference_sobolev_3}
+	\biggl\| r^{2t}\Delta_r(f_{2t}) - \sum_{h = t + 1}^{\floor{(s - 1)/2}}r^{2h}f_{2h}\biggr\|_{\Leb^2(\Vset_{(q + 1)r})} \leq c M r^{s}.
+	\end{equation}
+	On the other hand, when $2t = s - 2$ or $2t = s - 1$, it follows directly  from~\eqref{eqn:norm_expected_difference_operator} that
+	\begin{equation}
+	\label{pf:expected_difference_sobolev_4}
+	\Bigl\|r^{2t}\Delta_r(f_{2t})\Bigr\|_{\Leb^2(\Vset_{qr})} \leq c  r^{2t + (s - 2t)} \norm{f_{2t}}_{H^{s - 2t}(\Vset_{qr})} \leq c M r^{s}.
+	\end{equation}
+
+	Finally, noting that $\Delta_p^qf \in H^{s - 2q}(\Xset,M) \subset H^3(\Vset_{qr},M)$, we may again apply the inductive hypothesis and obtain
+	\begin{equation*}
+	\biggl\| r^{2q} I_{2e_1}^q \Delta_r(\Delta_p^qf) - r^{2(q + 1)} I_{2e_1}^{q + 1} \Delta_p^{q + 1}f - \sum_{t = 2}^{\floor{(s - 2q - 1)/2}} r^{2(t + q)}f_{2(t + q)}\biggr\|_{\Leb^2(V_{(q + 1)r})} \leq c M r^{2q + (s - 2q)}
+	\end{equation*}
+	or, again writing things more succinctly,
+	\begin{equation}
+	\label{pf:expected_difference_sobolev_5}
+	\biggl\| r^{2q} I_{2e_1}^q \Delta_r(\Delta_p^qf) - r^{2(q + 1)} I_{2e_1}^{q + 1} \Delta_p^{q + 1}f - \sum_{h = q + 2}^{\floor{(s - 1)/2}} r^{2h}f_{2h}\biggr\|_{\Leb^2(V_{(q + 1)r})} \leq c M r^{s}.
+	\end{equation}
+	The claim~\eqref{pf:expected_difference_sobolev_0} then follows from~\eqref{pf:expected_difference_sobolev_2}-\eqref{pf:expected_difference_sobolev_5} and the triangle inequality, completing the inductive step and thus the proof of Lemma~\ref{lem:expected_difference_sobolev}.
+\end{proof}
+
+\begin{proof}[\textbf{Proof of Lemma~\ref{lem:expected_difference_boundary}}]
+	When $q = 1$, we can relate the $\Leb^2$ norm of $\Delta_rf$ over $\Xset^r$ to the $\Leb^2$ norm of $f$ over $\Xset^{2r}$, as follows:
+	\begin{align*}
+	\Bigl\|\Delta_rf\Bigr\|_{\Leb^2(\Xset^r)}^2 & = \int_{\Xset^r} \biggl[\int_{\Xset} \bigl(f(y) - f(x)\bigr)K_r(x,y) p(y) \,dy \biggr]^2 \,dx \\
+	& \leq p_{\max}^2 \int_{\Xset^r} \biggl[\int_{\Xset} \bigl(\abs{f(y)} +  \abs{f(x)}\bigr)K_r(x,y) \,dy \biggr]^2 \,dx \\
+	& \overset{(i)}{\leq}  p_{\max}^2 K_{\max}^2 \int_{\Xset^r} \biggl[\int_{B(0,1) \cap (\Xset - x)/r} \bigl(\abs{f(zr + x)} +  \abs{f(x)}\bigr)\,dz \biggr]^2 \,dx \\
+	& \overset{(ii)}{\leq} 2 p_{\max}^2 K_{\max}^2 \int_{\Xset^r} \nu_d^2 \bigl(f(x)\bigr)^2 \,dx + 2 p_{\max}^2 K_{\max}^2 \nu_d \int_{\Xset^r}\biggl[\int_{B(0,1) \cap (\Xset - x)/r} \bigl(f(zr + x)\bigr)^2\,dz\biggr] \,dx \\
+	& \leq 2 p_{\max}^2 K_{\max}^2 \int_{\Xset^r} \nu_d^2 \bigl(f(x)\bigr)^2 \,dx + 2 p_{\max}^2 K_{\max}^2 \nu_d \int_{B(0,1)} \int_{\Xset^r} \bigl(f(zr + x)\bigr)^2\,dz \,dx \\
+	& \leq 4 p_{\max}^2 K_{\max}^2 \nu_d^2 \norm{f}_{\Leb^2(\Xset^{2r})}^2
+	\end{align*}
+	where $(i)$ follows from change of variables and $(ii)$ from Young's and Jensen's inequality.
+Reasoning by induction, we see that to prove the claim of Lemma~\ref{lem:expected_difference_boundary} it suffices to establish that
+	\begin{equation}
+	\label{pf:expected_difference_boundary}
+	\norm{f}_{\Leb^2(\Xset^{(q + 1)r})}^2 \leq c r^{2s} \norm{f}_{H^s(\Xset)}^2,
+	\end{equation}
+	and in Subsection~\ref{subsec:proof_sobolev_norm_boundary} we show that this inequality holds whenever $r > 0$ is sufficiently small.
+\end{proof}
+
+\begin{proof}[\textbf{Proof of Theorem~\ref{thm:graph_sobolev_seminorm}}]
+	From the representation~\eqref{eqn:graph_sobolev_seminorm_1}, we have that 
+	\begin{equation}
+	\label{pf:graph_sobolev_seminorm}
+	\Ebb\Bigl[f^T \Lap_{n,r}^s f\Bigr] = r^{ds}\sum_{i = 1}^{n} \sum_{k \in [n]^q} \sum_{\ell \in [n]^q} \Ebb\Bigl[\bigl(D_kf\bigr)(X_i) \cdot  \bigl(D_{\ell}f\bigr)(X_i)\Bigr].
+	\end{equation}
+	We divide into cases based on the cardinality of $\abs{k \cup \ell \cup i}$.
+If $\abs{k \cup \ell \cup i} = 2q + 1$, the intersection between any two of $k \in (n)^q$, $\ell \in (n)^q$, and $i \in [n]$ is empty.
+As a result, we can relate the expected product of difference operators to the square of the expected difference operator $\Delta_r^qf$ by applying the law of iterated expectation, resulting in
+	\begin{align*}
+	\Ebb\Bigl[\bigl(D_kf\bigr)(X_i) \cdot \bigl(D_{\ell}f\bigr)(X_i)\Bigr] & = \Ebb\Bigl[\bigl(\Delta_r^qf(X)\bigr)^2\Bigr] \\
+	& \leq p_{\max}^2 \norm{\Delta_r^qf}_{\Leb^2(\Xset)}^2 \\
+	& = p_{\max}^2 \Bigl\{\norm{\Delta_r^qf}_{\Leb^2(\Xset_{qr})}^2 + \norm{\Delta_r^qf}_{\Leb^2(\Xset^{qr})}^2\Bigr\} \\
+	& \leq p_{\max}^2 \Bigl\{\norm{\Delta_r^qf}_{\Leb^2(\Xset_{qr})}^2 + c M r^{2s}\Bigr\} 
+	\end{align*}
+	where the last inequality follows from Lemma~\ref{lem:expected_difference_boundary}.
+Since $f \in H^{s}(\Xset)$ for $s = 2(q - 1) + 2$, Lemma~\ref{lem:expected_difference_sobolev} allows us to upper bound $\norm{\Delta_r^{q - 1}f}_{\Leb^2(\Xset_{(q - 1)r})}$, which we use to obtain the upper bound
+	\begin{align}
+	\norm{\Delta_r^qf}_{\Leb^2(\Xset_{qr})} & = \norm{\Delta_r(\Delta_r^{q - 1}f)}_{\Leb^2(\Xset_{qr})} \nonumber \\
+	& \leq \Bigl\|\Delta_r\Bigl(\Delta_r^{q - 1}f - r^{2(q - 1)}I_{2e_1}^q\Delta_p^{q - 1}f\Bigr)\Bigr\|_{\Leb^2(\Xset_{qr})} +  r^{2(q - 1)}I_{2e_1}^q\Bigl\|\Delta_r\Bigl(\Delta_p^{q - 1}f\Bigr)\Bigr\|_{\Leb^2(\Xset_{qr})} \nonumber \\
+	& \leq c \Bigl\|\Delta_r^{q - 1}f - r^{2(q - 1)}I_{2e_1}^q\Delta_p^{q - 1}f\Bigr\|_{\Leb^2(\Xset_{(q - 1)r})} +  c r^{s}I_{2e_1}^q\Bigl\|\Delta_p^{q - 1}f\Bigr\|_{H^2(\Xset_{(q - 1)r})} \nonumber \\
+	& \leq c M r^s
+	\end{align}
+	with the second inequality following from~\eqref{eqn:norm_expected_difference_operator}.
+There are at most $O(n^{s + 1})$ such terms, hence the overall contribution of these terms to the expectation~\eqref{pf:graph_sobolev_seminorm} is $O(n^{s + 1}r^{s(d + 2)})M^2$.
+
+	Otherwise $\abs{k \cup \ell \cup i} = I < 2q + 1$.
+By Lemma~\ref{lem:expected_difference_lower_order_1} the total contribution of such terms inside the sum~\eqref{pf:graph_sobolev_seminorm} is $O(n^{I}r^{d(I - 1) + 2}) M^2$; by assumption $r \geq n^{-1/d}$, and therefore this increases with $I$.
+Taking $I = s$ to be the largest integer less than $s + 1$, the contribution of these terms to the sum is therefore $O(n^sr^{d(s - 1) + 2})M^2$, which in light of the restriction $r \geq n^{-1/(2(s - 1) + d)}$ is $O(n^{s+1}r^{s(d +2)})M^2$.
+\end{proof}
 
@@
 
-Let $\Delta_r$ be the expected weighted difference operator
-\begin{equation}
-\Delta_rf(x) := \Ebb\Bigl[\bigl(D_if\bigr)(x)\Bigr] = \int_{\Xset} \bigl(f(x') - f(x)\bigr) K_r(x',x) p(x') \,dx
-\end{equation}
-and define $\Delta_r^q = \Delta_r \circ \cdots \circ \Delta_r$, where the operator is composed a total of $q$ times. 
 \begin{lemma}
@@
 	\begin{equation*}
-	\Bigl\|\bigl(\Delta_r^qf\bigr)(x) - r^{2q} \sigma_K^{2q} \bigl(\Delta_{p}^qf\bigr)(x)\Bigr\| \leq C \cdot
+	\Bigl|\bigl(\Delta_r^qf\bigr)(x) - r^{2q} \sigma_K^{2q} \bigl(\Delta_{p}^qf\bigr)(x)\Bigr| \leq C \cdot
 	\begin{dcases}
@@
 	\end{equation*}
-	Finally, if $\abs{\alpha + \beta} = u \geq 4$ and $\abs{\beta} \leq u - 1$, then $f^{(\alpha)} \in C^{s - \abs{\alpha}}(\Xset,M) \subset C^{s- u}(\Vset_r,M)$ and  $p^{(\beta)} \in C^{(s - 1) - \abs{\beta}}(\Xset,M) \subset C^{s - u}(\Vset_r,M)$.
+	Finally, if $\abs{\alpha + \beta} = u \geq 4$ and $\abs{\beta} \leq u - 1$, then $f^{(\alpha)} \in C^{s - \abs{\alpha}}(\Xset,M) \subset C^{s- u}(\Vset_r,M)$\footnote{Again, technically speaking we mean that if $g \in C^{s - \abs{\alpha}}(\Xset,M)$ then $\restr{g}{\Vset_r} \in C^{s- u}(\Vset_r,M)$.} and  $p^{(\beta)} \in C^{(s - 1) - \abs{\beta}}(\Xset,M) \subset C^{s - u}(\Vset_r,M)$.
 Along with~\eqref{eqn:expected_difference_holder_2}, this verifies $f_u \in C^{s - u}(\Vset_r,cM)$ for a constant $c$ which does not depend on $f$ or $M$, and completes the proof of~\eqref{eqn:expected_difference_holder_0} in the base case.
@@
 
+\subsection{Proof of~\eqref{eqn:integrated_remainder_term_sobolev} and~\eqref{eqn:integrated_remainder_term_holder}}
+\label{subsec:proof_integrated_remainder_terms}
+
+First, we bound the $\Leb^2(V)$ norm of $\Ebb_{\alpha}[R_f^{\alpha}]$ by an integral of $R_f^{(\alpha)}(x,x')$ squared,
+\begin{align*}
+\int_{V} \Bigl[\Ebb_{\alpha}[R_f^{\alpha}]\bigr)(x)\Bigr]^2 \,dx & =  \int_{V} \biggl[\int_{\Vset} R_f^{(\alpha)}(x,x') (x' - x)^{\alpha} K_r(x,x') p(x') \,dx' \biggr]^2 \,dx \\
+& = r^{2(s + d)} \int_{V} \biggl[\int_{B(0,1)} R_f^{(\alpha)}(x,zr + x) (z)^{\alpha} K(\norm{z}) p(zr + x) \,dz \biggr]^2 \,dx \\
+& \leq \frac{1}{\nu_d}r^{2(s + d)} (p_{\max} K_{\max})^2 \int_{V} \int_{B(0,1)} \Bigl[R_f^{(\alpha)}(x,zr + x) \Bigr]^2  \,dz  \,dx \\
+& \leq \frac{1}{\nu_d}r^{2(s + d)} (p_{\max} K_{\max})^2 \int_{B(0,1)} \int_{V}  \Bigl[R_f^{(\alpha)}(x,zr + x) \Bigr]^2  \,dx  \,dz.
+\end{align*}
+Then by Jensen's inequality and then Fubini's Theorem,
+\begin{align*}
+\int_{V} \Bigl[R_f^{(\alpha)}(u,w + u)\Bigr]^2 \,du & = \int_{V} \biggl\{\int_{0}^{1} (1 - t)^{s} f^{(\alpha)}(u + tw) \,dt  \biggr\}^2 \,du \\ 
+& \leq \int_{V} \int_{0}^{1} \Bigl[f^{(\alpha)}(u + tw)\Bigr]^2 \,dt   \,du \\
+& = \int_{0}^{1} \int_{V} \Bigl[f^{(\alpha)}(u + tw)\Bigr]^2 \,du  \,dt \\
+& \leq \int_{0}^{1} \int_{U} \Bigl[f^{(\alpha)}(u)\Bigr]^2 \,du  \,dt \leq |f|_{H^s(U)}^2,
+\end{align*}
+where the penultimate inequality is where we invoke that $B(x,r) \subset U$ for every $x \in U$. This proves~\eqref{eqn:integrated_remainder_term_sobolev}.
+
+The proof of~\eqref{eqn:integrated_remainder_term_holder} follows along the same lines, replacing $\Leb^2$ norms by $\Leb^{\infty}$ norms where necessary.
+
+\subsection{Norms of $\Delta_r$.}
+\label{subsec:proof_boundedness_expected_difference_operator}
+
+We prove~\eqref{eqn:norm_expected_difference_operator} for the case $s = 0$, then $s = 1$, and finally $s = 2$.
+
+To start we assume $f \in H^0(V) = \Leb^2(V)$ for an open set $V \subseteq \Xset$. Then writing $\norm{\Delta_rf}_{\Leb^2(V_r)}^2$ as an integral,  
+\begin{align*}
+\int_{V_r} \biggl\{\frac{1}{r^d} \int_{\Xset} \bigl(f(x') - f(x)\bigr) K_r(x',x) p(x') \,dx'\biggr\}^2 \,dx & = \nu_d^2 \int_{V_r} \biggl\{\frac{1}{\nu_d r^d} \int_{B(x,r)} \bigl(f(x') - f(x)\bigr) K_r(x',x) p(x') \,dx'\biggr\}^2 \,dx \\
+& \leq \nu_d \int_{V_r} \frac{1}{r^d} \int_{B(x,r)} \Bigl\{\bigl(f(x') - f(x)\bigr) K_r(x',x) p(x') \,dx'\Bigr\}^2 \,dx \\
+& \leq \frac{2 \nu_d }{r^d} K_{\max}^2 p_{\max}^2 \int_{V_r} \int_{B(x,r)} f(x')^2 + f(x)^2 \,dx' \,dx \\
+& = 2 \nu_d K_{\max}^2 p_{\max}^2 \int_{V_r} \int_{B(0,1)} f(x + zr)^2 + f(x)^2 \,dz \,dx \\
+& \leq 4 \nu_d^2 K_{\max}^2 p_{\max}^2 \norm{f}_{\Leb^2(V)}^2.
+\end{align*}
+
+In the latter two cases---$s = 1$ and $s = 2$---we will assume that $f \in C^{\infty}(V)$, so that we may take Taylor expansions.
+This is without loss of generality: the class $C^{\infty}(V)$ is dense in $H^s(V)$, and consequently for any $f \in H^s(V)$ and $\delta > 0$ there exists $f_{\delta} \in C^{\infty}(V)$ such that $\norm{f - f_{\delta}}_{H^s(V)} \leq \delta$.
+Therefore if~\eqref{eqn:norm_expected_difference_operator} holds for all $C^{\infty}(V)$ functions, it also holds for all $H^s(V)$ functions; this follows from
+\begin{equation*}
+\norm{\Delta_rf}_{\Leb^2(V_r)} \leq \norm{\Delta_r(f_{\delta})}_{\Leb^2(V_r)} + \delta \leq cr^s \norm{f_{\delta}}_{H^s(V)} + \delta \leq \delta(cr^s + 1) + cr^s\norm{f}_{H^s(V)}
+\end{equation*}
+and taking the limit as $\delta \to 0$.
+
+When $s = 1$ the proof follows quite simply from~\eqref{eqn:integrated_remainder_term_sobolev},
+\begin{align*}
+\norm{\Delta_rf}_{\Leb^2(V_r)} & = \frac{1}{r^d}\Big\|\sum_{i = 1}^d \Ebb_{e_i}[R_f^{(e_i)}]\Bigr\|_{\Leb^2(V_r)} \\
+& \leq \frac{1}{r^d}\sum_{i = 1}^{d} \Big\| \Ebb_{e_i}[R_f^{(e_i)}]\Bigr\|_{\Leb^2(V_r)} \\
+& \leq d p_{\max} K_{\max} r |f|_{H^1(V)}.
+\end{align*}
+
+Finally, when $s = 2$ and $p \in C^1(\Xset)$, a simple computation verifies that
+\begin{equation*}
+\bigl(\Delta_rf\bigr)(x) = \sum_{i = 1}^{d} \sum_{j = 1}^{d} \Bigl(\Ibb_{e_i + e_j}[R_p^{(e_j)}]\Bigr)(x) \cdot f^{(e_i)}(x) + \sum_{\abs{\alpha} = 2}  \Bigl(\Ebb_{\alpha}[R_f^{(\alpha)}]\Bigr)(x).
+\end{equation*}
+Therefore, using~\eqref{eqn:integrated_remainder_term_sobolev}-\eqref{eqn:product_rule_sobolev}, we obtain
+\begin{align*}
+\bigl\|\Delta_rf\bigr\|_{\Leb^2(V_r)} & \leq \frac{1}{r^d} \biggl\{ \Bigl\|\sum_{i = 1}^{d} \sum_{j = 1}^{d} \Ibb_{e_i + e_j}[R_p^{(e_j)}] \cdot f^{(e_i)} \Bigr\|_{\Leb^2(V_r)} + \Bigl\|\sum_{\abs{\alpha} = 2} \Ebb_{\alpha}[R_f^{(\alpha)}]\Bigr\|_{\Leb^2(V_r)} \biggr\} \\
+& \leq \frac{C}{r^d} \biggl\{ \sum_{i = 1}^{d} \sum_{j = 1}^{d} \Bigl\|\Ibb_{e_i + e_j}[R_p^{(e_j)}]\Bigr\|_{\Leb^{\infty}(V_r)} \cdot \Bigl\|f^{(e_i)} \Bigr\|_{\Leb^2(V_r)} + \sum_{\abs{\alpha} = 2} \Bigl\|\Ebb_{\alpha}[R_f^{(\alpha)}]\Bigr\|_{\Leb^2(V_r)} \biggr\} \\
+& \leq c r^2 \norm{f}_{H^2(V)}.
+\end{align*}
+
+\subsection{Proof of~\eqref{pf:expected_difference_boundary}}
+\label{subsec:proof_sobolev_norm_boundary}
+We will prove that the claim holds for all $g \in C_c^{\infty}(\Xset)$, whence the density of $C_c^{\infty}(\Xset)$ in $H_0^s(\Xset)$ implies that it holds for all $f$ in the latter class as well.
+
+Fix $x_0 \in \partial \Xset$, and let $Q_d(x_0,r)$ be the $d$-dimensional cube centered at $x_0$ of side length $r$. We will show that for all sufficiently small $r > 0$,
+\begin{equation}
+\norm{g}_{\Xset \cap \Leb^2(Q_d(x_0,r))} \leq c r^s \norm{g}_{H^s(Q_d(x_0,r))}
+\end{equation}
+The Lemma then follows by taking a finite covering of the boundary $\partial \Xset$ -- possible since $\Xset$ is assumed to be bounded -- in a similar manner to e.g.
+Theorem 18.1 of \textcolor{red}{(Leoni)} or Theorem 1 in 5.5 of \textcolor{red}{(Evans)}.
+
+We begin by straightening the boundary.
+Since $\Xset$ has a Lipschitz boundary, for any $x_0 \in \partial \Xset$ there exists a rigid motion $\Phi: \Rd \to \Rd$ with $T(x_0) = 0$, a Lipschitz continuous function $\gamma: \Reals^{d-1} \to \Reals^d$, and a radius $r_0' > 0$ such that, setting $y = \Phi(x)$ and writing $y = (y',y_d)$, we have that for all $r < r_0$,
+\begin{equation*}
+\Phi(\Xset \cap Q(x_0,r)) = \bigl\{y \in Q(0,r): y_d > \gamma(y')\bigr\}.
+\end{equation*}
+Fixing $0 < r < r_0/[\mathrm{Lip}(\gamma)\sqrt{d}]$, we have that $\gamma(y') > -r$ for all $y' \in Q_{d - 1}(x_0,r)$; writing $\wt{g}(y) = g(T^{-1}(y))$, taking a Taylor expansion of $\wt{g}(y)$ around $\wt{g}((y',\gamma(y')))$ thus yields
+\begin{align*}
+\wt{g}(y) & = \sum_{\ell = 1}^{s - 1} g^{(\ell e_d)}\bigl((y',\gamma(y'))\bigr) (y_d - \gamma(y'))^{\ell} + \int_{\gamma(y')}^{y_d} \wt{g}^{(se_d)}\bigl((y',he_d)\bigr) h^{s - 1}\,dh \\
+& = \int_{\gamma(y')}^{y_d} \wt{g}^{(se_d)}\bigl((y',he_d)\bigr) h^{s - 1}\,dh
+\end{align*}
+where the second equality follows from the assumption $g \in C_c^{\infty}(\Xset)$. We now analyze the $\Leb^2$ norm over $Q_d(x_0,r)$,
+\begin{align*}
+\norm{g}_{\Xset \cap \Leb^2(Q_d(x_0,r))}^2 & = \norm{\wt{g}}_{\Leb^2(\Phi(\Xset \cap Q(0,r)))}^2 \\
+& = \int_{Q_{d - 1}(0,r)} \int_{\gamma(y')}^{r} \biggl[\int_{\gamma(y')}^{y_d} \wt{g}^{(se_d)}\bigl((y',he_d)\bigr)h^{s - 1} \,dh\biggr]^2 \,dy_d \,dy' \\
+& \overset{(i)}{\leq} (2r)^{2(s - 1)} \int_{Q_{d - 1}(0,r)} \int_{\gamma(y')}^{r} \bigl(y_d - \gamma(y')\bigr)^2 \biggl[\frac{1}{y_d - \gamma(y')}\int_{\gamma(y')}^{y_d} \wt{g}^{(se_d)}\bigl((y',he_d)\bigr) \,dh\biggr]^2 \,dy_d \,dy' \\
+& \overset{(ii)}{\leq}  (2r)^{2(s - 1)} \int_{Q_{d - 1}(0,r)} \int_{\gamma(y')}^{r} \bigl(y_d - \gamma(y')\bigr) \int_{\gamma(y')}^{y_d} \biggl[\wt{g}^{(se_d)}\bigl((y',he_d)\bigr)\biggr]^2 \,dh \,dy_d \,dy' \\
+& \overset{(iii)}{\leq}  (2r)^{2s - 1} \int_{Q_{d - 1}(0,r)} \int_{\gamma(y')}^{r} \int_{\gamma(y')}^{r} \biggl[\wt{g}^{(se_d)}\bigl((y',he_d)\bigr)\biggr]^2 \,dh \,dy_d \,dy' \\
+& \leq  (2r)^{2s} \int_{Q_{d - 1}(0,r)}  \int_{\gamma(y')}^{r} \biggl[\wt{g}^{(se_d)}\bigl((y',he_d)\bigr)\biggr]^2 \,dh  \,dy' \\
+& \overset{(iv)}{\leq} (2r)^{2s} \int_{Q(x_0,r)} \bigl[g^{(se_d)}(x)\bigr]^2 \,dx \leq (2r)^{2s} \norm{g}_{H^s(B(x_0,r))}^2
+\end{align*} 
+where $(i)$ follows since $0 < y_d - \gamma(y') < r - \gamma(y') < 2r$, $(ii)$ follows by Jensen's inequality, $(iii)$ follows since $y_d < r$, and $(iv)$ follows from a change of variables.
+This completes the proof.
+
 \end{document}

### 2020-09-15 notes/graph_regression2.tex
Draft of graph regression paper nearly complete.

--- 2020-08-26 notes/graph_regression2.tex
@@
 \newcommand{\seq}[1]{\set{#1}_{n \in \N}}
-\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
 \newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
@@
 \newcommand{\Cov}{\mathrm{Cov}}
-\newcommand{\Xsetiam}{\mathrm{diam}}
+\newcommand{\diam}{\mathrm{diam}}
 
@@
 \newcommand{\SM}{\mathrm{SM}}
-\newcommand{\OS}{\mathrm{OS}}
+\newcommand{\OS}{\mathrm{FS}}
 \newcommand{\PLS}{\mathrm{PLS}}
@@
 \section{Introduction}
-
-In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ with support $\Xset \subset \Rd$, and
+\label{sec:introduction}
+
+In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Rd$, and
 \begin{equation}
@@
 where $\kappa \in \{1,...,n\}$ is a tuning parameter.
-(Here $\dotp{u}{v}_n = \frac{1}{n}\sum_{i = 1}^{n} u_i v_i$ for vectors $u,v \in \Reals^n$.) The Laplacian smoothing estimator $\wt{f}_{\LS}$ \citep{smola2003} is given by
+(Here $\dotp{u}{v}_n = \frac{1}{n}\sum_{i = 1}^{n} u_i v_i$ for vectors $u,v \in \Reals^n$.) The Laplacian smoothing estimator $\wt{f}_{\LS}$ \citep{smola2003} is a penalized least squares estimator, given by
 \begin{equation}
@@
 \end{equation}
-can be used to test whether or not $f_0 = 0$.
-
-\begin{itemize}
-	\item \textbf{Motivate graph Laplacians.} There already exist a vast number of methods for both the nonparametric testing and, especially, the nonparametric estimation problems.
+are in turn reasonable statistics to test whether or not $f_0 = 0$.
+
+There already exist a vast number of methods for both the nonparametric testing and, especially, the nonparametric estimation problems.
 In fact, as we will see in Section~\ref{sec:problem_setup_and_background} some of these methods are actually quite analogous to the graph Laplacian estimation and testing methods we have just introduced.
 Nevertheless, methods such as~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} have a deserved place in the nonparametric regression toolbox.
-Their advantages include:
+Graph-based methods such as these have many advantages: they are easy to implement, fast and stable to compute, generalizable to non-standard data---for instance, text or images, or really any data modality on which it is possible to define a kernel--- and non-standard learning paradigms---such as the semi-supervised setting---and last but not least, have enjoyed widespread practical success.
+
+For these reasons, a substantial body of work has arisen assessing the statistical properties of graph Laplacian based methods.
+Taken as a whole, this work views the graph Laplacian matrix as an estimator of a weighted continuum Laplace operator $\Delta_P$ (see equations~\eqref{eqn:laplace_operator} and~\eqref{eqn:laplace_operator_weighted} in Section~\ref{sec:problem_setup_and_background}), and establishes that it is in fact a \emph{consistent} estimator of $\Delta$, in various respects.
+For instance,  \citet{belkin03,belkin05} show that when $\Pbb$ is the uniform distribution over $\Xset$, $\Lap_{n,r}f \to \Delta f$ for sufficiently smooth functions $f$, \citet{lafon04,hein05} extend the statement to hold when $\Pbb$ is non-uniform, \citet{singer06} sharpens the convergence rates, and \cite{gine06} establish a CLT.
+On the other hand, \citet{belkin07} prove that when $\Pbb$ is the uniform distribution over $\mathcal{X}$, the eigenvalues and eigenvectors of $\Lap_{n,r}$ converge to those of $\Delta$ when $r_n \to 0$ as $n \to \infty$, \citet{garciatrillos18} extend this result to hold for general $\Pbb$, and \cite{calder2019} obtain sharper rates at which $r_n$ may converge to $0$.
+
+These developments notwithstanding, the following fundamental theoretical question remains unanswered: are graph-based methods for nonparametric regression not merely consistent but in fact \emph{optimal}, in the minimax sense?
+In this paper, we confirm that they are.
+We show that when the regression function $f_0$ is suitably smooth---by which we mean that it belongs to a first-order Sobolev ball; roughly speaking, that it possesses weak derivative $Df$ bounded in $L^2(\Xset)$ norm---both~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are with high probability asymptotically rate-optimal estimators, and that both~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} are asymptotically rate-optimal test statistics.
+We also show that under appropriate conditions, these estimators and test statistics can leverage additional assumed structure, either on the regression function $f_0$ or distribution $P$.
+Our fundamental message is a reassuring one: the practitioner who wishes to use neighborhood graph procedures, for any of the various reasons outline above, may do so without fear that they are sacrificing statistical optimality in the process.
+
 \begin{itemize}
-		\item Simplicity.
-		\item Computational efficiency and stability.
-		\item Domain and density adaptivity. Easily generalizable. 
-		\item Practical success.
-	\end{itemize}
-	\item \textbf{Our contributions, short version.} For these reasons, there is a  substantial body of work assessing the statistical properties of graph Laplacian based methods.
-\textcolor{red}{(TODO): Include citations.} In this paper, we adopt a different perspective, and analyze~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} through the minimax lens.
-We show that when the regression function $f_0$ is suitably smooth (by which we mean that it belongs to a first-order Sobolev ball; roughly speaking, that it possesses weak derivative $Df$ bounded in $L^2(\Xset)$ norm) both~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are minimax optimal estimators, and both~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} are minimax optimal test statistics.
-We also show that under appropriate conditions, these estimators and test statistics can leverage additional assumed structure, either on the regression function $f_0$ or distribution $P$.
+	\item \textbf{Motivate graph Laplacians.}
+
+	\item \textbf{Literature review.} 
 	\item \textbf{Organization.}
@@
 \begin{itemize}
-	\item We write $\mathbb{Z}^d$ for $\mathbb{Z} \otimes \cdots \mathbb{Z}$, where the tensor product is taken a total of $d$ times.
-	\item We denote by $\Leb^2(\Xset)$ be the set of square integrable functions $f:\Xset \to \Reals$.
+	\item We write $\mathbb{Z}^d = \mathbb{Z} \otimes \cdots \otimes \mathbb{Z}$ (the tensor product is taken a total of $d$ times); the sequence space $\ell^2(\mathbb{Z}^d)$ contains those vectors $\theta: \mathbb{Z}^d \to \Reals$ for which $\sum_{k \in \mathbb{Z}^d} \theta_k^2 < \infty$.
+	\item We denote by $\Leb^2(\Xset)$ the set of square integrable functions $f:\Xset \to \Reals$. \textcolor{red}{Holder}.
 	\item For functions $f,g$ which map from $\Xset$ to $\Reals$, we write $\dotp{f}{g}_n := \frac{1}{n}\sum_{i = 1}^{n} f(X_i) g(X_i)$, which we refer to as the \emph{empirical inner product}, and $\norm{f}_n^2 = \dotp{f}{f}_n$, which we refer to as the (squared) \emph{empirical norm}.
 For vectors $v \in \Reals^n$, we write $\norm{v}_p^p := \sum_{i = 1}^{n} \abs{v_i}^p$.
-	\item For a graph $G$, we use $\Lap_G$ for the Laplacian of $G$.
+	\item For a graph $G$, we use $\Lap_G$ to denote the Laplacian of $G$.
 The matrix $\Lap_G$ is a square, symmetric, positive semi-definite matrix, with $n$ eigenvalues $\lambda_k(G)$ and eigenvector $v_k(G)$ pairs, defined according to
@@
 	\end{equation*}
-	and ordered so that $0 \leq \lambda_1(G) \leq \ldots \leq \lambda_n(G)$. For convenience, we write $\Lap_{n,r}$ instead of $\Lap_{G_{n,r}}$. 
+	and ordered so that $0 \leq \lambda_1(G) \leq \ldots \leq \lambda_n(G)$. For convenience, when dealing with the neighborhood graph $G_{n,r}$ we denote $\Lap_{n,r} = \Lap_{G_{n,r}}$. 
+	\item When convenient we write $\mathbf{X} = (X_1,\ldots,X_n)$. 
 \end{itemize}
@@
 Before we get to our main results, we will review some of the relevant literature on minimax optimal nonparametric estimation and testing of Sobolev functions.
-We will also briefly recall some (of the many) estimators and tests which achieve these rates: in particular the \emph{truncated series} and \emph{smoothing spline} methods.
+We will also briefly recall some (of the many) estimators and tests which achieve these rates: in particular the \emph{Fourier series projection} and \emph{smoothing spline} methods.
 As we will see, Laplacian eigenmaps and Laplacian smoothing can be seen as graph-based, i.e.
-discrete, analogues to these ``continuuum'' methods.
+discrete, counterparts to these ``continuum'' methods.
+
+We start by giving a (very) brief exposition of Sobolev spaces, with an emphasis on the material relevant in the statistical context.
 
@@
 
-\begin{itemize}
-	\item \textbf{Sobolev spaces.}  Roughly speaking, Sobolev spaces contain functions $f \in \Leb^p(\Xset)$ with derivatives which themselves belong to $\Leb^p(\Xset)$.
-More formally, for given integers $s, p > 0$, the Sobolev space $W^{s,p}(\Xset)$ consists of all functions $f \in \Leb^p(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^p(\Xset)$.
+Roughly speaking, Sobolev spaces contain functions $f \in \Leb^p(\Xset)$ with derivatives which themselves belong to $\Leb^p(\Xset)$.
+More formally, for given integers $s$ and $p > 0$, the Sobolev space $W^{s,p}(\Xset)$ consists of all functions $f \in \Leb^p(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^p(\Xset)$.
 The Sobolev-$\{s,p\}$ seminorm $\seminorm{f}_{W^{s,p}(\Xset)}$ and norm $\norm{f}_{W^{s,p(\Xset)}}$ are then given by
@@
 and for a given $M > 0$, the corresponding ball is $W^{s,p}(\Xset, M) = \set{f: \norm{f}_{W^{s,p}(\Xset)} \leq M}$.
-In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space, and we adopt the usual convention of writing $H^s(\Xset) = W^{s,p}(\Xset)$ and $H^s(\Xset,M) = W^{s,2}(\Xset,M)$.
-We will confine our attention to this special case hereafter.
-	\item \textbf{Connections to sequence spaces.} The balls $H^s(\Xset,M)$ are closely connected with the \emph{Sobolev ellipsoids} $\Theta_s(M) \subseteq \ell^2(\mathbb{Z}^d)$, defined as
-	\begin{equation*}
-	\Theta_s(M) = \Bigl\{\theta \in \ell_2(\mathbb{Z}^d): \sum_{k \in \mathbb{Z}^d} \theta_k^2 \abs{k}^{2s} < M\Bigr\}
-	\end{equation*}
-	Let us assume that the domain $\Xset = [0,1]^d$. Letting be $\set{\phi_k}_{k \in \mathbb{Z}^d}$ be the tensor product trigonometric polynomial basis of $\Leb^2(\Xset)$, the space
+In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space, and we adopt the usual convention of writing $H^s(\Xset) = W^{s,p}(\Xset)$ and $H^s(\Xset,M) = W^{s,2}(\Xset,M)$; we will confine our attention to this case hereafter.
+One special fact worth mentioning: when $\Xset$ is sufficiently smooth, for all Sobolev spaces $H^s(\Xset)$ the Holder space $C^{\infty}(\Xset)$ is dense in $H^s(\Xset)$.
+Said differently, for each integer $s$, the Sobolev space $H^s(\Xset)$ is simply the completion of $C^{\infty}(\Xset)$ in the norm $\norm{\cdot}_{H^s(\Xset)}$.
+
+Periodic functions $f \in H^s(\Xset)$ admit a representation in terms of Fourier coefficients. Define the \emph{Sobolev ellipsoids} $\Theta_s(M) \subseteq \ell^2(\mathbb{Z}^d)$ to be
+\begin{equation*}
+\Theta_s(M) = \Bigl\{\theta \in \ell_2(\mathbb{Z}^d): \sum_{k \in \mathbb{Z}^d} \theta_k^2 \abs{k}^{2s} \leq M^2\Bigr\}.
+\end{equation*}
+To see the connection between $H^s(M)$ and $\Theta_s(M)$, for the moment let us fix the domain $\Xset = [0,1]^d$, and write $\set{\phi_k}_{k \in \mathbb{Z}^d}$ for the tensor product trigonometric polynomial basis of $\Leb^2(\Xset)$.
+Then the periodic Sobolev balls
 \begin{equation*}
@@
 \end{equation*}
-	satisfies $H_{\textrm{per}}^s(\Xset,M) \subseteq H^s(\Xset,M')$ for a sufficiently large value of $M_1 > 0$.
-On the other hand, any $1$-periodic function $f \in W^{s,p}(\Xset,M)$ also belongs to $H_{\textrm{per}}^s(\Xset,M_0)$ for a sufficiently large value of $M_0 > 0$.
-	\item \textbf{Zero-trace Sobolev spaces.} We will also be interested in the \emph{zero-trace} Sobolev spaces.
-\end{itemize}
-
-\subsection{Orthogonal series and smoothing splines.}
+satisfy $H_{\textrm{per}}^s(\Xset,M) \subseteq H^s(\Xset,M')$ for a sufficiently large value of $M_1 > 0$.
+On the other hand, any $1$-periodic function $f \in H^s(\Xset,M)$ also belongs to $H_{\textrm{per}}^s(\Xset,M_0)$ for a sufficiently large value of $M_0 > 0$.
+
+We will also be interested in the \emph{zero-trace} Sobolev spaces, which impose more severe boundary conditions than periodicity.
+Letting $\partial \Xset$ denote the boundary of $\Xset$, informally one can think of the order-$s$ zero-trace Sobolev space as the set of functions $f \in H^s(\Xset)$ for which ``$f(x) = 0$ for all $x \in \partial\Xset$''.
+However, since functions $f \in H^s(\Xset)$ are defined only up to sets of measure zero, such a definition is not really meaningful.
+Instead, letting $C_c^{\infty}(\Xset)$ consist of those functions in $C^{\infty}(\Xset)$ which are compactly supported in $\Xset$, we formally define the order-$s$ zero-trace Sobolev space $H_0^{s}(\Xset)$ as the completion of $C_c^{\infty}(\Xset)$ in the $H^{s}(\Xset)$ norm.
+When the domain $\Xset = [0,1]^d$, it holds that $H_0^{s}(\Xset) \subseteq H_{\textrm{per}}^{s}(\Xset)$.
+\textcolor{red}{Justify this statement with a reference.}
+
+\textcolor{red}{(TODO 1)}: Refer also to the Sobolev embedding theorem, for at least three reasons: (1) to facilitate later discussion of why smoothing splines are ill-posed when $2s < d$; (2) to make explicit why the statistical problem isn't ill-posed when $2s < d$ in the random design setting; (3) to make clear why the testing problem is hopeless when $4s < d$.
+
+\subsection{Nonparametric regression: Fourier series and smoothing splines.}
 \label{subsec:continuum_methods}
-When the regression function is assumed to belong to a Sobolev space, two approaches particularly well-suited for nonparametric regression are orthogonal series \textcolor{red}{projection}, and smoothing splines.
-The orthogonal series estimator $\wh{f}_{\OS}$ is given by
+When the regression function $f_0$ is assumed to belong to a Sobolev space, two approaches particularly well-suited for nonparametric regression are orthogonal projection onto the Fourier basis, and smoothing splines.
+In the former, we estimate the Fourier coefficients $\wh{\theta}_k := \dotp{Y}{\phi_k}_n$, with the resultant function estimate $\wh{f}_{\OS}$ defined as
 \begin{equation}
 \label{eqn:orthogonal_series}
-\wh{f}_{\OS} = \sum_{k: \abs{k} \leq \kappa} \dotp{Y}{\phi_k}_n \phi_k .
-\end{equation}
-where $\kappa \geq 0$ is a hyperparameter, and we recall that $\{\phi_k\}_{k \in \mathbb{Z}^d}$ is the tensor product trigonometric polynomial basis of $\Leb^2(\Xset)$.
-The smoothing spline estimator $\wt{f}_{\SM}$ is a variational estimator, defined as the solution to the following optimization problem,
+\wh{f}_{\OS} = \sum_{k: \abs{k} \leq \kappa} \wh{\theta}_k \phi_k
+\end{equation}
+for some $\kappa \geq 0$. In contrast, the smoothing spline estimator $\wt{f}_{\SM}$ is a variational estimator, defined as the solution to the following optimization problem,
 \begin{equation}
 \label{eqn:smoothing_spline}
-\wt{f}_{\SM} = \argmin_{f} \sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \seminorm{f}_{H^{s}(\Xset)}^2
-\end{equation}
-where the minimum is taken over all $f \in H^s(\Xset)$.
-
-\begin{itemize}
-	\item \textbf{Connecting orthogonal series and Laplacian eigenmaps estimators.}The orthogonal series estimator $\wh{f}_{\OS}$ is closely connected with the Laplacian eigenmaps estimator $\wh{f}_{\LE}$.
-At a high-level, one can go from~\eqref{eqn:orthogonal_series} to~\eqref{eqn:laplacian_eigenmaps} simply by substituting $\{v_k\}_{k \in [n]}$ for $\{\phi_k\}_{k \in \mathbb{Z}^d}$.
-More deeply, just as the vectors $v_k$ are eigenvectors of the graph Laplacian $\Lap_{n,r}$, so too are the functions $\phi_k$ eigenfunctions of a particular differential operator $\Delta: C^2(\Xset) \to \Leb^2(\Xset)$, known as the \emph{Laplace} operator and formally defined as
+\wt{f}_{\SM} = \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \seminorm{f}_{H^{s}(\Xset)}^2\biggr\}
+\end{equation}
+with the minimum taken over all $f \in H^s(\Xset)$.
+
+The orthogonal series estimator $\wh{f}_{\OS}$ is closely connected with the Laplacian eigenmaps estimator $\wh{f}_{\LE}$.
+Just as $v_k(G_{n,r})$ are eigenvectors of the graph Laplacian $\Lap_{n,r}$, so too are $\phi_k$ eigenfunctions of a particular differential operator $\Delta: C^2(\Xset) \to \Leb^2(\Xset)$, known as the \emph{Laplace} operator and formally defined as
 \begin{equation}
 \label{eqn:laplace_operator}
-	\Delta g := -\mathrm{div}(\nabla g).
-	\end{equation}
-	In fact, it has been shown in \textcolor{red}{(references)} that when the radius $r \to 0$ as $n \to \infty$, the graph Laplacian $\Lap_{n,r}$ converges in an appropriate sense to a weighted version $\Delta_{P}$ of $\Delta$, with
-	\begin{equation}
-	\Delta_{P} g := -\frac{1}{p}\mathrm{div}(p^2 \nabla g).
-	\end{equation}
-	Along similar lines, \textcolor{red}{(references)} have established that for any fixed $k \geq 0$, as $n \to \infty$ the eigenvector $v_k$ will converge, again in an appropriate sense, to an eigenfunction $\phi_k$ of $\Delta_{P}$.
-The upshot is that when $P$ is the uniform measure, for an appropriate sequence of radii $r(n) \to 0$ and any fixed $\kappa \geq 0$ the estimator $\wh{f}_{OS}$ is in fact the limit of $\wh{f}_{\LE}$, in the sense that
+\Delta g := -\mathrm{div}(\nabla g),
+\end{equation}
+where $\mathrm{div}$ stands for the divergence, and $\nabla$ for the gradient.
+As mentioned previously, when the radius $r \to 0$ as $n \to \infty$ at an appropriate rate, the graph Laplacian $\Lap_{n,r}$ in fact converges to a weighted Laplace operator $\Delta_P$, given by
+\begin{equation}
+\label{eqn:laplace_operator_weighted}
+\Delta_{P} g := -\frac{1}{p}\mathrm{div}(p^2 \nabla g),
+\end{equation}
+with the eigenvectors $v_k(G_{n,r})$ converging in an appropriate sense to eigenfunctions of $\Delta_{P}$.
+The upshot is that when $P$ is the uniform measure---and thus the eigenfunctions of $\Delta_P = \Delta$ are in fact the trigonometric basis functions $\phi_k$---for an appropriate sequence of radii $r(n) \to 0$ and any fixed $\kappa \geq 0$ the estimator $\wh{f}_{\OS}$ is in fact the limit of $\wh{f}_{\LE}$, in the sense that
 \begin{equation}
@@
 for each $i \in \mathbb{N}$.
-	\item 
-	\textbf{Connecting smoothing spline and Laplacian smoothing estimators.} In a similar vein, the smoothing spline estimator $\wh{f}_{\SM}$ has deep connections with the Laplacian smoothing estimator $\wh{f}_{\LS}$.
-Using the definition~\eqref{eqn:laplace_operator}, recursively applying integration by parts, and then invoking the aforementioned convergence properties of $\Lap_{n,r}$, it can be shown that \textcolor{red}{references} that for any $f \in C^{2s}(\Xset)$,
+
+By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can also tie the smoothing spline estimator $\wh{f}_{\SM}$ to the Laplacian smoothing estimator $\wh{f}_{\LS}$.
+Using the definition~\eqref{eqn:laplace_operator}, recursively applying integration by parts, and then invoking the aforementioned convergence properties of $\Lap_{n,r}$, it can be shown that for any $f \in C^{2s}(\Xset)$,
 \begin{equation}
 \label{eqn:pointwise_seminorm_convergence}
-	\abs{f}_{H^s(\Xset)} = \int_{\Xset} \bigl(\Delta^sf\bigr)(x) f(x) \,dx = \lim_{n \to \infty} f^T \Lap_{n,r}^s f.
-	\end{equation}
-	By applying appropriation concentration inequalities, one can extend~\eqref{eqn:pointwise_seminorm_convergence} to hold uniformly over all $f \in H^{2s}(\Xset)$.
-While this does not imply a formal statement about the convergence of $\wh{f}_{\LS}$ to $\wh{f}_{SM}$ as in~\eqref{eqn:laplacian_eigenmaps_consistency}---note the difference between the domain of minimization in~\eqref{eqn:smoothing_spline} and the regularity required for~\eqref{eqn:pointwise_seminorm_convergence}---it nevertheless makes clear that the estimators $\wh{f}_{\LS}$ and $\wh{f}_{\SM}$ are linked.
-	\item \textbf{Domain and density adaptivity.} \textcolor{red}{(TODO)}
-\end{itemize}
+\abs{f}_{H^s(\Xset)} = \int_{\Xset} \bigl(\Delta_P^sf\bigr)(x) f(x) \,dx = \lim_{n \to \infty} \frac{1}{n^{s + 1}r^{s(d + 2)}}f^T \Lap_{n,r}^s f.
+\end{equation}
+(See \citet{zhou11} for an argument along these lines.)
+While this does not imply a formal statement about the convergence of $\wh{f}_{\LS}$ to $\wh{f}_{SM}$ akin to~\eqref{eqn:laplacian_eigenmaps_consistency}---note, for instance, the difference between the domain of minimization in~\eqref{eqn:smoothing_spline} and the regularity required for~\eqref{eqn:pointwise_seminorm_convergence}---it nevertheless makes the connection between $\wh{f}_{\LS}$ and $\wh{f}_{\SM}$ obvious.
 
@@
 \label{subsec:minimax_rates}
-One reason all of this is intriguing is that, as we mentioned previously, the continuum estimators $\wh{f}_{\OS}$ and $\wh{f}_{\SM}$---and test statistics based on functionals of these estimators--- have strong statistical properties.
-In particular, these estimators and tests are known to have error rates within a constant factor of (minimax) optimal.
+One reason all of this is intriguing is that, as mentioned previously, the continuum estimators $\wh{f}_{\OS}$ and $\wh{f}_{\SM}$---and test statistics based on functionals of these estimators--- have strong statistical properties.
+In particular, these estimators and tests are known to have error rates within a constant factor of minimax optimal.
 
@@
 \label{eqn:sobolev_space_estimation_minimax_rate}
-\inf_{\wh{f}} \sup_{f_0 \in \mc{H}^s(M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d \geq 0$,}
-\end{equation}
-where the infimum is over all measurable maps $\wh{f}(X_1,Y_1,\ldots,X_n,Y_n)$, and the function class $\mc{H}^s(M)$ is one of $H^s(\Xset,M)$, $H_{\mathrm{per}}^s(\Xset,M)$ or $H_0^s(\Xset,M)$.
-Orthogonal series and smoothing spline estimators are two (of the many) estimators which certify the upper bound in~\eqref{eqn:sobolev_space_estimation_minimax_rate}.
-Concretely, if we are willing to assume
+\inf_{\wh{f}} \sup_{f_0 \in \mc{H}^s(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d \geq 0$,}
+\end{equation}
+where the infimum is over all measurable maps $\wh{f}$, and the function class $\mc{H}^s(\Xset,M)$ is any one of $H^s(\Xset,M)$, $H_{\mathrm{per}}^s(\Xset,M)$ or $H_0^s(\Xset,M)$.
+Fourier series and smoothing spline estimators are two (of the many) estimators which certify the upper bound in~\eqref{eqn:sobolev_space_estimation_minimax_rate}.
+In the former case, if we assume that
 \begin{enumerate}[label=(P\arabic*)]
@@
 	\label{asmp:uniform_density}
-	the distribution $P$ is the uniform probability measure defined on domain $\Xset = [0,1]^d$.
+	the distribution $P$ is the uniform probability measure on the unit cube $\Xset = [0,1]^d$.
 \end{enumerate}
-then the orthogonal series estimator is minimax optimal over the Sobolev ellipsoids $H_{\mathrm{per}}^s(M)$, meaning
+then $\wh{f}_{\OS}$ is minimax optimal over the Sobolev ellipsoids $H_{\mathrm{per}}^s(\Xset, M)$, meaning
 \begin{equation*}
@@
 \end{equation*}
-An advantage of $\wt{f}_{\SM}$ over $\wh{f}_{\OS}$ is that it requires less stringent assumptions on the design density in order to be minimax optimal.
+An advantage of $\wt{f}_{\SM}$ over $\wh{f}_{\OS}$ is that it requires weaker assumptions on the design density in order to be minimax optimal.
 Instead of Assumption~\ref{asmp:uniform_density}, we will merely assume that the density is bounded above and below.
@@
 	\label{asmp:bounded_density}
-	The distribution $P$ is defined on domain $\Xset = [0,1]^d$, and admits a density $p$ with respect to the uniform (Lebesgue) measure on $[0,1]^d$.
 	The density $p$ is bounded away from $0$ and $\infty$, that is
@@
 	\end{equation*}
-	for all $x \in [0,1]^d$.
+	for all $x \in \Xset$.
 \end{enumerate}
-The assumption~\ref{asmp:bounded_density} is itself a standard but significant assumption; we will return to discuss it further at \textcolor{red}{a later point}.
+The assumption~\ref{asmp:bounded_density} is itself a standard but significant assumption, but clearly it is much less stringent than~\ref{asmp:uniform_density}: it does not specify that either the density $p$ or domain $\Xset$ be known.
 Assuming~\ref{asmp:bounded_density}, the estimator $\wt{f}_{\SM}$ satisfies
@@
 \end{equation*}
-The restriction $2s > d$ is a severe limitation of smoothing splines relative to  orthogonal series.
-It is worth emphasizing that this is a fundamental defect of the smoothing spline estimator, as opposed to any flaw in the theory; when $2s \leq d$, the optimization problem~\eqref{eqn:smoothing_spline} is not even well-defined!
-\textcolor{red}{(TODO):} succinctly explain why.
+The restriction $2s > d$ is a severe limitation of smoothing splines relative to  orthogonal series estimators.
+It is worth emphasizing that this is a fundamental defect of smoothing splines, as opposed to any flaw in the theory; when $2s \leq d$, the optimization problem~\eqref{eqn:smoothing_spline} is not even well-defined!
+
+\textcolor{red}{(TODO 2):} succinctly explain why, by referring to the Sobolev Embedding Theorem.
 
@@
 
-\begin{itemize}
-	\item \textbf{Minimax hypothesis testing.} In the nonparametric regression hypothesis testing problem, we ask for a test function (formally, a Borel measurable map $\phi: \mc{X}^n \times \Reals^n \to \{0,1\}$) which can distinguish between the null and alternative hypotheses
+In the nonparametric regression goodness-of-fit testing problem, we ask for a test function (formally, a Borel measurable map $\phi: \mc{X}^n \times \Reals^n \to \{0,1\}$) which can distinguish between the null and alternative hypotheses
 \begin{equation}
@@
 \end{equation}
-	We measure the performance of a test $\phi$ by its Type I error $\Ebb_0(1 - \phi)$, and Type II error $\mathbb{E}_{f_0}(\phi)$.
-It is well known that no test $\phi$ with controlled Type I error---say, $\Ebb_0(\phi) \leq 1/b$ for some constant $b > 1$---
-	can have controlled Type II error---$\Ebb_{f_0}(\phi) \leq 1/b$--- for every $f_0 \neq 0$.
+We measure the performance of a test $\phi$ by its Type I error $\Ebb_0[1 - \phi]$, and Type II error $\mathbb{E}_{f_0}[\phi]$.
+It is well known that no test $\phi$ with controlled Type I error---say, $\Ebb_0[\phi] \leq 1/b$ for some constant $b > 1$---
+can have controlled Type II error---$\Ebb_{f_0}[\phi] \leq 1/b$--- for every $f_0 \neq 0$.
 As a result, it is standard to impose two kinds of restrictions on the alternative hypothesis class $\mathbf{H}_a$.
-The first is a smoothness condition on $f_0$: specifically, we will assume that $f_0$ belongs to a Sobolev ball $\mc{H}^s(M)$.
-The latter is a separation condition, enforcing a minimum distance between $f_0$ and $0$: specifically, we will insist that $\norm{f_0  - 0}_{\Leb^2(\Xset)} > \epsilon$ for some $\epsilon > 0$.
-	\item\textbf{Minimax testing rates over Sobolev classes.} Once smoothness and separation conditions are established, the following becomes a natural question to ask: what is the smallest value of $\epsilon$ for which there exists a test $\phi$ with both small Type I error, and small Type II error over all functions $f_0$ which are smooth and separated from $0$ by at least $\epsilon$?
+The first is a smoothness condition on $f_0$: in our case, we will assume that $f_0$ belongs to a Sobolev ball $\mc{H}^s(\Xset,M)$.
+The latter is a separation condition, enforcing a minimum distance between $f_0$ and $0$: in our case, we will insist that $\norm{f_0  - 0}_{\Leb^2(\Xset)} > \epsilon$ for some $\epsilon > 0$.
+
+Once smoothness and separation conditions are established, the following becomes a natural question to ask: what is the smallest value of $\epsilon$ for which there exists a test $\phi$ with both small Type I error, and small Type II error over all functions $f_0$ which are smooth and separated from $0$ by at least $\epsilon$?
 This smallest separation $\epsilon$ is known as the critical radius, and in the case of the Sobolev classes it is given by
-	\begin{equation*}
-	\epsilon_n^{\star}\bigl(\mc{H}^s(M)\bigr):= \inf\biggl\{\epsilon: \inf_{\phi} \Bigl\{\Ebb_0(\phi) +  \sup_{\substack{f_0 \in \mc{H}^s(M) \\ \norm{f_0}_{\Leb^2(\Xset)} \geq \epsilon}} \Ebb_{f_0}[1 - \phi]\Bigr\} \leq \frac{2}{b} \biggr\} \asymp n^{-2s/(4s + d)}
-	\end{equation*}
-	for any $4s > d$, and $\mc{H}^s(M)$ one of $H^s(\Xset,M)$, $H_{\mathrm{per}}^s(\Xset,M)$ or $H_0^s(\Xset,M)$. On the other hand, when $4s \leq d$ the minimax rate is dimension-free,
-	\begin{equation*}
-	\epsilon_n^{\star}\Bigl(\mc{H}^s(M)\Bigr) \asymp n^{-1/4},
-	\end{equation*}
-	revealing a distinction between the estimation and goodness-of-fit testing problems.
-	\item \textbf{Error rates of orthogonal series and smoothing spline-based tests.} 
+\begin{equation}
+\label{eqn:sobolev_space_testing_critical_radius}
+\epsilon_n^{\star}\Bigl(\mc{H}^s(\Xset,M)\Bigr):= \inf\biggl\{\epsilon: \inf_{\phi} \Bigl\{\Ebb_0[\phi] +  \sup_{\substack{f_0 \in \mc{H}^s(\Xset,M), \\ \norm{f_0}_{\Leb^2(\Xset)} \geq \epsilon}} \Ebb_{f_0}[1 - \phi]\Bigr\} \leq \frac{2}{b} \biggr\} \asymp n^{-2s/(4s + d)}
+\end{equation}
+for any $4s > d$, and $\mc{H}^s(\Xset,M)$ any one of $H^s(\Xset,M)$, $H_{\mathrm{per}}^s(\Xset,M)$ or $H_0^s(\Xset,M)$.
 Just as $\wh{f}_{\OS}$ and $\wh{f}_{\SM}$ achieve minimax optimal estimation rates, tests using the plug-in statistics
@@
 \end{equation*}
-	are minimax optimal for the goodness-of-fit testing problem over Sobolev balls (see \citet{ingster2009} for analysis of the former statistic, and \citet{liu2019} for analysis of the latter).
-\end{itemize}
+are minimax optimal for the goodness-of-fit testing problem over Sobolev balls; see \citet{ingster2009} for analysis of the former statistic, and \citet{liu2019} for analysis of the latter (when $2s > d$).
+
+When $4s \leq d$, the Sobolev classes $H^s(\Xset,M)$ are too irregular for any test statistic to accurately discern whether $\norm{f_0}_{\Leb^2(\Xset)} = 0$; in particular, some functions in these classes will not have a bounded $4$th moment.
+However, we will show in Section~\ref{sec:minimax_optimal_graph_Laplacian_methods} that upon explicitly assuming a bounded 4th moment, one obtains the dimension-free minimax rate
+\begin{equation*}
+\epsilon_n^{\star}\Bigl(\Leb^4(\mc{X},M)\Bigr) \asymp n^{-1/4},
+\end{equation*}
+revealing a distinction between the estimation and goodness-of-fit testing problems. This extends the results of \cite{guerre02}, who show the analogous result over Holder balls.
+
+\paragraph{Manifold hypothesis.}
+
+Suppose that $\Xset$ satisfies the manifold hypothesis, meaning informally that it has intrinsic dimension $m < d$ (see~\ref{asmp:domain_manifold} in Section~\ref{sec:graph_sobolev_classes} for a formal definition).
+In this context, both~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} now hold with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
+What's more, suitable adaptations of the estimators $\wh{f}_{\OS}$ and $\wt{f}_{\SM}$, and the test statistics $T_{\OS}$ and $T_{\SM}$, achieve these faster rates.
+We will not further review the details here, except to emphasize one point: in this context, in order for either the smoothing spline or Fourier series approaches to be optimal, the domain $\Xset$ must be known; as we will see, our graph-based procedures will be optimal without this requirement.
+We will return to this topic, and comment more generally on how different methods and analyzes make use of different assumptions on the design distribution $P$, in Section~\ref{sec:discussion}.
 
@@
 
-\begin{itemize}
-	\item \textbf{Departure from what's already known.} At a high level, the connections between the graph-based methods~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} and the continuum methods outlined in Subsection~\ref{subsec:continuum_methods} give reason to believe that the strong statistical properties of the latter are shared by the former.
+At a high level, the connections between the graph-based methods~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} and the continuum methods outlined in Subsection~\ref{subsec:continuum_methods} give us reason to hope that the strong statistical properties of the latter are shared by the former.
 Making this rough statement precise is the content of this paper.
 In particular, we will prove the following results:
+
 \begin{itemize}
 	\item \textbf{Minimax optimal estimation.}
-		We show that with high probability, $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2s/(2s + d)}$ over all $f_0 \in H^1(\Xset;M)$.
-The same conclusions hold with respect to $\wt{f}_{\LS}$ when $d \leq 4$.
+	With high probability, $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2/(2 + d)}$ over all $f_0 \in H^1(\Xset;M)$. The same conclusions hold with respect to $\wt{f}_{\LS}$ when $d \leq 4$. 
 	\item \textbf{Minimax optimal testing.}
-		We show that with high probability, a test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in H^1(\Xset;M)$ satisfies $\norm{f_0}_{\Leb^2(\Xset)} \gtrsim \max\{n^{-2s/(4s + d)},n^{-1/4}\}$.
-\textcolor{red}{The same conclusions hold with respect to $\wt{f}_{LS}$.}
-		\item \textbf{Taking advantage of higher-order smoothness conditions.} Under appropriate boundary conditions on the function classes $H^s(\Xset)$ and additional smoothness conditions on the density $p$, when $s > 1$  each of the aforementioned conclusions hold for certain values of $d$.
+	A test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in H^1(\Xset;M)$ satisfies $\norm{f_0}_{\Leb^2(\Xset)} \gtrsim n^{-2/(4 + d)}$ and $d < 4$.
+The same conclusions hold with respect to $T_{LS}$.
+	\item \textbf{Taking advantage of higher-order smoothness conditions.} Under appropriate boundary conditions on the function classes $H^s(\Xset)$ and additional smoothness conditions on the density $p$, with high probability the error $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2s/(2s + d)}$ over all $f_0 \in H^s(\Xset;M)$.
+Additionally, a test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in H^s(\Xset;M)$ satisfies $\norm{f_0}_{\Leb^2(\Xset)} \gtrsim \max\{n^{-2s/(4s + d)}\}$ and $d < 4$.
 	\item \textbf{Manifold adaptivity.}
@@
 \end{itemize}
+
+\section{Graph Sobolev classes}
+\label{sec:graph_sobolev_classes}
+In this section, we introduce graph Sobolev balls, which as we will see are discrete (graph-based) analogues to continuum Sobolev balls, and establish some relations between the discrete and continuum classes.
+For an integer $s > 0$ and some $M > 0$, define the $s$th-order graph Sobolev ball to be
+\begin{equation}
+\label{eqn:graph_sobolev_ball}
+H^s(G_{n,r},M) = \Bigl\{f \in \Reals^n: f^T \Lap_{n,r}^s f \leq M^2\Bigr\}
+\end{equation}
+where we recall that the \emph{$s$th-order graph Sobolev seminorm} $f^T \Lap_{{n,r}}^s f$ is precisely the penalty term in~\eqref{eqn:laplacian_smoothing}.
+
+To understand the relevance of the graph Sobolev balls to the analysis of our graph-based procedures, let us return for a moment to the continuum setting.
+The statistical analysis of continuous-time methods such as Fourier series projection and smoothing splines rests on two facts about the continuum classes $\mc{H}^s(\Xset)$: first, an \textit{a priori} assumption that the regression function $f_0$ belongs to an appropriate Sobolev ball $\mc{H}(\Xset,M)$, and second that the asymptotic decay of the eigenvalues of the continuum Laplace operator $\Delta$ obey Weyl's law, meaning $\lambda_k(\Delta) \asymp \abs{k}^{2s}$ for $k \in \mathbb{Z}^d$.
+In this section, we will show that both these statements have rough analogues in the graph setting.
+Roughly speaking, we will show first that for functions $f$ belonging to $\mc{H}^s(\Xset, M)$, the evaluations $f = (f(X_1),\ldots,f(X_n)) \in \Reals^n$ satisfy the upper bound
+\begin{equation}
+\label{eqn:graph_sobolev_seminorm_asymptotic}
+f^T \Lap_{n,r}^s f \lesssim M(n^{s + 1}r^{s(d + 2)})
+\end{equation}
+and second that the eigenvalues $\lambda_k(G_{n,r})$ satisfy the asymptotic scaling
+\begin{equation}
+\label{eqn:neighborhood_graph_eigenvalue_asymptotic}
+\bigl[\lambda_k(G_{n,r})\bigr]^s = \lambda_k(\Lap_{{n,r}}^s) \asymp k^{2s/d} (n^s r^{s(d + 2)}).
+\end{equation}
+Both statements will hold with high probability.
+Additionally, when we assume $\Xset$ is a manifold of dimension $m$, they will hold with $d$ replaced by $m$.
+These results elucidate why our graph-based procedures ``work'' as well as their continuum counterparts---in the sense of being statistically rate-optimal over (continuum) Sobolev classes.
+They also clarify when our analysis does and does not apply---in terms of various orders of smoothness $s$ and dimensions $d$ (or $m$).
+
+\subsection{Graph Sobolev seminorm}
+\label{subsec:graph_sobolev_seminorm}
+We will state and prove three different results involving the graph Sobolev semi-norm $f_0^T \Lap_{n,r}^s f_0$.
+In each case, we will derive that a version of the asymptotic relation~\eqref{eqn:graph_sobolev_seminorm_asymptotic} holds with high probability, whenever $r \to 0$ sufficiently slowly as $n \to \infty$.
+The difference between the three results will be the assumptions we make, and the resulting rate of permissible decay for the radius $r$.
+
+\subsubsection{First-order graph Sobolev semi-norm.}
+\label{subsec:first_order_graph_sobolev_seminorm}
+We begin by upper bounding $f_0^T \Lap_{n,r} f_0$ under the assumption $f_0 \in H^1(\Xset,M)$.
+We separate this out from the analysis of general $s$ because the regularity conditions we require when $s = 1$ will be meaningfully weaker than when $s > 1$.
+Specifically, we will assume the kernel $K$ and domain $\Xset$ satisfy conditions~\ref{asmp:kernel} and~\ref{asmp:domain}, respectively.
+\begin{enumerate}[label=(K\arabic*)]
+	\item
+	\label{asmp:kernel}
+	$K:[0,\infty) \to [0,\infty)$ is supported on $[0,1]$, and its restriction to $[0,1]$ is Lipschitz. Additionally, it is normalized so that
+	\begin{equation*}
+	\int_{\Reals^d} K\bigl(\norm{z}\bigr) \,dz = 1.
+	\end{equation*}
+\end{enumerate}
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{2}
+	\item 
+	\label{asmp:domain}
+	\textcolor{red}{(TODO): Explain what a Lipschitz domain is.} The density $p$ is supported on a Lipschitz domain $\Xset$.
+\end{enumerate}
+The former is a relatively mild assumption: the choice of kernel is under the user's control, and moreover~\ref{asmp:kernel} covers many (though certainly not all) common kernel choices.
+The latter is somewhat more restrictive, but an important point is that we do not require that the domain be \textit{a priori} known in order to formulate our estimators and tests; we will elaborate on this point in Section~\ref{sec:discussion}.
+In Lemma~\ref{lem:first_order_graph_sobolev_seminorm}, we prove that under these conditions, if the density $p$ is merely upper bounded then~\eqref{eqn:graph_sobolev_seminorm_asymptotic} holds with respect to the first-order graph Sobolev semi-norm.
+\begin{lemma}
+	\label{lem:first_order_graph_sobolev_seminorm}
+	Assume~\ref{asmp:kernel} and~\ref{asmp:domain}, and additionally that $p(x) \leq p_{\max}$ for all $x \in \Xset$.
+Then, for any $f \in H^1(\Xset,M)$, there exists a constant $C_1 > 0$ which depends only on $\Xset$, $d$, $p_{\max}$ and $K$ such that
+	\begin{equation}
+	\label{eqn:first_order_graph_sobolev_seminorm}
+	f^T \Lap_{n,r} f \leq \frac{C_1}{\delta} n^2 r^{d + 2} M^2
+	\end{equation}
+	with probability at least $1 - \delta$.
+\end{lemma}
+The proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm}, as with all of our proofs, is contained in the Appendix.
+There are a few remarks worth making regarding Lemma~\ref{lem:first_order_graph_sobolev_seminorm}, but they all apply equally well to our analysis of $f^T\Lap_{{n,r}}^sf$ for $s > 1$, and so will wait to make them.
+
+\subsubsection{Higher-order graph Sobolev semi-norm}
+\label{subsec:higher_order_graph_sobolev_seminorm}
+When $s > 1$, in order to obtain the desired upper bound~\eqref{eqn:graph_sobolev_seminorm_asymptotic} on the asymptotic growth of the graph Sobolev semi-norm we must make two kinds of additional assumptions: the first on the behavior of $f_0$ near the boundary of $\Xset$, and the second on the smoothness of the density function $p$.
+\begin{theorem}
+	\label{thm:graph_sobolev_seminorm}
+	Fix some integer $s > 1$.
+Assume~\ref{asmp:kernel} and~\ref{asmp:domain}, and additionally suppose that the density $p \in C^{s - 1}(\Xset,p_{\max})$ and that $f \in H_0^s(\Xset,M)$.
+Let $r = r(n)$ be a sequence of connectivity radii satisfying $r(n) \to 0$ and $r(n) \gtrsim n^{-1/(2(s - 1) +d)}$ as $n \to \infty$.
+Then for all $n$ sufficiently large, there exists a constant $C_2 > 0$ which depends only on $\Xset$, $d$, $p_{\max}$ and $K$ such that
+	\begin{equation}
+	\label{eqn:graph_sobolev_seminorm}
+	f^T \Lap_{n,r}^s f \leq \frac{C_2}{\delta} n^{s + 1} r^{s(d + 2)}M^2
+	\end{equation}
+	with probability at least $1 - \delta$.
+\end{theorem}
+Now, we make some remarks:
+\begin{itemize}
+	\item 
+	The scaling (in $n$ and $r$) on the right hand side of~\eqref{eqn:graph_sobolev_seminorm} will be familiar to the informed reader.
+For a sufficiently regular function $f \in C^3(\Xset,M)$ and density $p \in C^1(\Xset,p_{\max})$, it is known \textcolor{red}{(references)} that $\frac{1}{nr^{d + 2}}\Lap_{n,r}f \to \Delta_pf$ for $n^{-1/(2 + d)}$.
+A recursive argument shows that when $f \in C^{2s + 1}(\Xset)$, $p \in C^{2s}(\Xset)$, and $r$ is sufficiently large,
+	\begin{equation}
+	\label{eqn:graph_sobolev_seminorm_heuristic}
+	\Bigl(\frac{1}{nr^{d + 2}}\Lap_{n,r}\Bigr)^s f \to \Delta_p^sf \Longrightarrow \frac{1}{n^{s + 1} r^{s(d + 2)}} f^T \Lap_{n,r}^s f \to \int_{\Xset} f(x) \cdot  \Delta_p^sf(x) p(x) \,dx \leq c M^2.
+	\end{equation}
+
+	\textcolor{red}{(TODO 1)}: Figure out the right scaling of $r$ for this heuristic argument to go through.
+
+	In fact, the actual proof of Theorem~\ref{thm:graph_sobolev_seminorm} is rather more involved.
+Among other subtleties we highlight the mismatch, in the heuristic argument just given, between the number of smooth derivatives $f$ possesses ($2s + 1$) and the degree of the Laplacian operator $\Lap_{{n,r}}^s$ $(s)$; this is unsuitable for our purposes.
+We use a more careful analysis to prove Theorem~\eqref{thm:graph_sobolev_seminorm}, but note well that~\eqref{eqn:graph_sobolev_seminorm} is only an upper bound rather than a statement about consistency.
+This is the price we pay for insisting that the assumed number of smooth derivatives match the degree of the Laplacian; fortunately an upper bound will be sufficient for our purposes in Section~\ref{sec:minimax_optimal_graph_Laplacian_methods}.
+
+	\item In~\eqref{eqn:first_order_graph_sobolev_seminorm} and~\eqref{eqn:graph_sobolev_seminorm}, we state the upper bound as a function of the Sobolev norm $\norm{f}_{H^s(\Xset)}$ of $f$, rather than the Sobolev seminorm $|f|_{H^s(\Xset)}$.
+The derivations in~\eqref{eqn:graph_sobolev_seminorm_heuristic} might leave us some hope that it suffices to assume a bounded seminorm.
+Indeed, in Lemma~\ref{lem:first_order_graph_sobolev_seminorm}, we may relax this---that is, we may replace $M$ by $|f|_{H^1(\Xset)}$---because the graph Laplacian $\Lap_{n,r}$ annihilates the constant part of any function $f$.
+In the latter case, we may not replace $M$ by $|f|_{H^s(\Xset)}$.
+
+	\textcolor{red}{(TODO 2)}: Explain why.
+
+	\textcolor{red}{(TODO 3)}: Check with Ryan and Siva--- or if not them, Google--- to figure out whether boundedness in seminorm plus zero-trace condition implies boundedness in norm.
+Intuitively, it feels like it does.
+
+	\textcolor{red}{(TODO 4)}: Go back and check whether Fourier series projection requires boundedness in norm, or only in semi-norm.
+	\item We note also that~\eqref{eqn:first_order_graph_sobolev_seminorm} and~\eqref{eqn:graph_sobolev_seminorm}---which each hold with probability $1 - \delta$---are proportional to $1/\delta$, as opposed to $\log(1/\delta)$ as is typical of exponential concentration.
+The reason: the Sobolev assumption $f \in H^s(\Xset,M)$ only grants us control over the second moment of $f$ and its derivatives, and so we are afforded only the relatively weak concentration implied by Markov's inequality.
+Under higher-order moment control on $f$ and its derivatives---for instance if $f \in C^s(\Xset,M)$---we can get sharper upper bounds.
+\end{itemize}
+
+Finally, we note that while the particular requirement in Theorem~\ref{thm:graph_sobolev_seminorm} that $f$ be zero-trace may be a tad overkill, it is essential that $f$ display some type of periodicity for the upper bound~\eqref{eqn:graph_sobolev_seminorm} to hold.
+
+\textcolor{red}{(TODO 5)}: Elaborate on this: (a) reference developments in the literature showing that the eigenvectors of $\Lap_{n,r}$ satisfy Neumann boundary conditions in the limit $n \to \infty, r \to 0$, and (b) state, as a Lemma, that there exists a function $g$ which is not zero-trace, and for which the upper bound~\eqref{eqn:graph_sobolev_seminorm} is not satisfied.
+
+\subsubsection{Graph Sobolev semi-norm under manifold assumption.}
+
+As mentioned in the introduction, an attractive quality of the neighborhood graph is that it automatically captures the geometry of $P$ and $\Xset$.
+We now show one manifestation of this phenomenon: the graph Sobolev semi-norm automatically adapts to the intrinsic dimension of $\Xset$.
+Formally, we make the following assumption on $\Xset$.
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{3}
+	\item 
+	\label{asmp:domain_manifold}
+	The measure $P$ is supported on a compact, connected, smooth submanifold $\Xset$ of $\Reals^d$; the domain $\Xset$ is of fixed dimension $1 \leq m \leq d$ and without boundary.
+Moreover, $P$ has a density $p$ with respect to the volume form of $\Xset$ which is bounded away from $0$ and $\infty$,
+	\begin{equation*}
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty
+	\end{equation*}
+	for all $x \in \Xset$.
+\end{enumerate}
+In Theorem~\ref{thm:manifold_graph_sobolev_seminorm}, the scaling of the graph Sobolev semi-norm now depends on the intrinsic dimension $m$, rather than the ambient dimension $d$.
+\begin{theorem}
+	\label{thm:manifold_graph_sobolev_seminorm}
+	Fix some integer $1 \leq s \leq 3$.
+Assume~\ref{asmp:kernel},~\ref{asmp:domain_manifold} and additionally suppose that $p \in C^{s - 1}(\Xset,p_{\max})$ and $f \in C^s(\Xset,M)$.
+Let $r = r(n)$ be a sequence of connectivity radii satisfying $r(n) \to 0$ and $r(n) \gtrsim n^{-1/(2(s - 1) + m)}$ as $n \to \infty$.
+Then for all $n$ sufficiently large, there exists a constant $C_3 > 0$ which depends only on $\Xset, m$ and $p$ such that
+	\begin{equation*}
+	f^T \Lap_{n,r}^s f \leq \frac{C_3}{\delta}n^{s + 1}r^{s(m + 2)}M^2
+	\end{equation*}
+	with probability at least $1 - \delta$.
+\end{theorem}
+\begin{itemize}
+	\item \textcolor{red}{(TODO 6)}: Explain why we require $s \leq 3$.
+	\item \textcolor{red}{(TODO 7)}: See if we can replace the Holder assumption $f \in C^s$ with the Sobolev assumption $f \in H^s$.
+\end{itemize}
+
+\subsection{Graph Laplacian eigenvalues}
+
+Lemma~\ref{lem:neighborhood_eigenvalue} establishes a scaling rate which holds for all eigenvalues $\lambda_{k}(G_{n,r})$, assuming appropriate regularity conditions on $P$.
+\begin{lemma}
+	\label{lem:neighborhood_eigenvalue}
+	Suppose the distribution $P$ satisfies~\ref{asmp:bounded_density} and~\ref{asmp:domain}, and the kernel $K$ satisfies~\ref{asmp:kernel}.
+Let $r = r(n)$ be a sequence of connectivity radii satisfying $r(n) \to 0$ as $n \to \infty$.
+Then for all $n$ sufficiently large, there exist constants $c_4$ and $C_4$ which depend only on $p_{\min}, p_{\max},d$ and $\Xset$, such that the following statement
+	\begin{equation}
+	\label{eqn:neighborhood_eigenvalue}
+	c_4 \cdot \min\Bigl\{nr^{d + 2}k^{2/d}, nr^d\Bigr\} \leq \lambda_k(G_{n,r}) \leq C_4 nr^{d + 2}k^{2/d}~~\textrm{for all $2 \leq k \leq n$,}
+	\end{equation}	
+	holds with probability at least $1 - C_4 r^{-d} \exp\bigl\{- c_4 n r^d\bigr\}$.
+\end{lemma}
+In Lemma~\ref{lem:neighborhood_eigenvalue_manifold}, we show that under the manifold assumption, these scalings hold but with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
+\begin{lemma}
+	\label{lem:neighborhood_eigenvalue_manifold}
+	Suppose the distribution $P$ satisfies~\ref{asmp:domain_manifold} and the kernel $K$ satisfies~\ref{asmp:kernel}.
+Let $r = r(n)$ be a sequence of connectivity radii satisfying $r(n) \to 0$ as $n \to \infty$.
+Then there exist constants $c_5$ and $C_5$ which depend only on $p_{\min}, p_{\max}$,$m$ and $\Xset$, such that the following statement
+	\begin{equation}
+	\label{eqn:neighborhood_eigenvalue_manifold}
+	 c_5 \cdot \min\Bigl\{nr^{m + 2}k^{2/m}, nr^m\Bigr\} \leq \lambda_k(G_{n,r}) \leq C_5 nr^{m + 2}k^{2/m}~~\textrm{for all $2 \leq k \leq n$,}
+	\end{equation}	
+	holds with probability at least $1 - C_5 r^{-m} \exp\bigl\{- c_5 n r^m\bigr\}$.
+\end{lemma}
+
+The second term in the lower bound of~\eqref{eqn:neighborhood_eigenvalue} and~\eqref{eqn:neighborhood_eigenvalue_manifold} is an artifact of the well known inequality $\lambda_n(G) \leq 2 \max_{i = 1,\ldots,n} \mathbf{D}_{ii}$---in words, that the maximum eigenvalue of the Laplacian can never be more than twice the maximum degree of the graph.
+To obtain optimal rates for our Laplacian eigenmaps estimator and test statistic, we will need the eigenvalues $\lambda_k(G_{n,r})$ to asymptotically scale like~\eqref{eqn:neighborhood_graph_eigenvalue_asymptotic} for particular values of $k$.
+Doing some algebra, we deduce in Corollary~\ref{cor:neighborhood_eigenvalue} various conditions on $k$ and $r$ under which the graph eigenvalue scales at least as fast as is desired, i.e.
+$\lambda_k(G_{n,r}) \gtrsim n r^{d + 2}k^{2/d}$.
+
+\begin{corollary}
+	\label{cor:neighborhood_eigenvalue}
+	Suppose the distribution $P$ satisfies~\ref{asmp:bounded_density} and~\ref{asmp:domain}, and the kernel $K$ satisfies~\ref{asmp:kernel}. Then each of the following statements:
+	\begin{itemize}
+		\item For $k = M^{2d/(2s + d)} n^{d/(2s + d)}$ and any $r \leq M^{-2/(2s + d)} n^{-1/(2s + d)}$, 
+		\begin{equation}
+		\label{eqn:neighborhood_eigenvalue_1}
+		\lambda_{k}(G_{n,r}) \geq c_4 \cdot nr^{d+2}{k}^{2/d}
+		\end{equation}
+		\item For $k = M^{4d/(4s + d)}n^{2d/(4s + d)}$ and any $r \leq M^{-4/(4s + d)} n^{-2/(4s + d)}$,
+		\begin{equation}
+		\label{eqn:neighborhood_eigenvalue_2}
+		\lambda_{k}(G_{n,r}) \geq c_4 \cdot nr^{d+2}{k}^{2/d}
+		\end{equation}
+	\end{itemize}
+	hold with probability at least $1 - C_4 r^{-d} \exp\{- c_4 n r^d\}$.
+
+	If instead the distribution $P$ satisfies~\ref{asmp:domain_manifold}, then each of the previous statements hold with $d$ replaced by $m$, and with the constants $C_4$ and $c_4$ replaced by $C_5$ and $c_5$, respectively.
+\end{corollary}
+
+Some remarks:
+\begin{itemize}
+	\item We will use inequality~\eqref{eqn:neighborhood_eigenvalue_1} to upper bound the bias of the Laplacian eigenmaps estimator $\wh{f}_{\LE}$, and inequality~\eqref{eqn:neighborhood_eigenvalue_2} to upper bound the bias of the Laplacian eigenmaps test statistic $T_{\LE}$.
+	\item Let us briefly summarize the various restrictions on the asymptotic scaling of $r$.
+Corollary~\ref{cor:neighborhood_eigenvalue} tells us that the desired eigenvalue scalings~\eqref{eqn:neighborhood_graph_eigenvalue_asymptotic} kick in only when the neighborhood graph radius $r$ is sufficiently small.
+On the other hand, in order for the probabilistic bounds of this corollary to be non-facile, the neighborhood graph radius must be sufficiently large, roughly speaking $r \gtrsim (\log n/n)^{1/d}$ (or in the manifold case $r \gtrsim (\log n/n)^{1/m}$).
+What's more, the conclusions of Theorem~\ref{thm:graph_sobolev_seminorm} hold only when neighborhood graph radius $r$ is sufficiently large.
+Consequently, our results in Section~\ref{sec:minimax_optimal_graph_Laplacian_methods} regarding statistical optimality of Laplacian eigenmaps will hold only for the values of $r$ which satisfy all these restrictions, when such values exist.
+	\item For Laplacian smoothing, in a rough sense the story is the same: to prove rate-optimal upper bounds, we will still want $\lambda_k(G_{n,r})$ to scale according to~\eqref{eqn:neighborhood_graph_eigenvalue_asymptotic}; to establish such a scaling we will require the neighborhood graph radius $r$ be small, but not so small that the probabilistic bounds are facile; and there will be attendant consequences for our theory in Section~\ref{sec:minimax_optimal_graph_Laplacian_methods}.
+However the technical details are a bit fussier than for Laplacian eigenmaps, and we consign them to the appendix.
 \end{itemize}
@@
 
-Let us start by assuming $f_0 \in H^1(P,M)$.
-In this section, we will establish that the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and the resulting test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test}, achieve the standard minimax optimal estimation and hypothesis testing rates, respectively.
-
-\subsection{Minimax optimal estimation}
-
-Under appropriate conditions, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves minimax rates over all $f_0 \in H^1(P,M)$ with high probability.
-The conditions we impose will be on the neighborhood graph $G_{n,r}$.
-As the construction of the graph is in the user's control, these assumptions are therefore mild.
-\begin{enumerate}[label=(K\arabic*)]
+We now turn to formalizing the main conclusions of this work: that in various settings and under various conditions, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and the resulting test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test}, are asymptotically minimax-rate optimal.
+We divide these conclusions into three parts.
+First, we will show that when $f_0$ itself is smooth---measured in the Sobolev $H^1(\Xset)$ norm---all of our graph-based procedures are minimax rate-optimal under minimal conditions.
+Second, we will show that when $f_0$ additionally possesses smooth (partial) derivatives, some of our graph-based procedures can---under certain conditions---provably achieve the faster minimax rates.
+Third, we will show that when the density $p$ is supported on a low-dimensional manifold our graph-based procedures again achieve faster rates than in the full-dimensional case, thereby confirming their domain-adaptive properties.
+
+Unless otherwise stated, for all results in this Section we will always assume that the density $p$ satisfies~\ref{asmp:bounded_density}, the domain $\Xset$ satisfies~\ref{asmp:domain}, and the kernel $K$ satisfies~\ref{asmp:kernel}.
+Additionally, within our Theorem statements we use $c$ and $C$ to denote constants which possibly differ from Theorem to Theorem,  and which may depend on $K$, $d$, $p_{\min}$, $p_{\max}$, $\Xset$ and $s$, but do not depend on $f_0$, the radius of the Sobolev ball $M$, or the sample size $n$.
+
+\subsection{First-order smoothness.}
+\label{subsec:minimax_first_order_smoothness}
+Let us start by assuming $f_0 \in H^1(\Xset,M)$, and proving upper bounds on the risk of our various graph-based procedures.
+Pleasingly, these results will go through with essentially no additional regularity conditions---additional, that is, to the standard~\ref{asmp:bounded_density} and~\ref{asmp:domain}---on either the regression function $f_0$ or the density $p$.
+The story will also be complete: we will obtain the ``right'' rates for all scales---that is, all values of $s$ and $d$---at which we would expect to get such rates.
+
+\subsubsection{Estimation}
+\label{subsubsec:first_order_smoothness_estimation}
+
+Under appropriate conditions, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves, up to constants, the minimax risk $n^{-2/(2 + d)}$ over all $f_0 \in H^1(\Xset,M)$.
+To prove this, we will invoke Lemma~\ref{lem:first_order_graph_sobolev_seminorm} and Corollary~\ref{lem:neighborhood_eigenvalue} to suitably upper bound the bias of $\wh{f}_{\LE}$: in order to do so, we must restrict the scaling of neighborhood graph radii $r(n)$.
+\begin{enumerate}[label=(R\arabic*)]
 	\item 
-	\label{asmp:kernel_form}
-	The kernel function $K(z) = \1\Bigl\{\norm{z}_2/r \leq 1 \Bigr\}$.
-	\item 
-	\label{asmp:kernel_radius_estimation}
-	The radii $r = r(n)$ satisfies $r(n) \to 0$ as $n \to \infty$, and additionally
-	\begin{equation*}
-	n^{-1/d}\leq r(n) \leq M^{2/(2 + d)}n^{-1/(2 + d)}
+	\label{asmp:le_kernel_radius_estimation}
+	The sequence of radii $r = r(n)$ satisfies
+	\begin{equation*}
+	\biggl(\frac{\log n}{n}\biggr)^{1/d} \lesssim r(n) \lesssim M^{-2/(2 + d)}n^{-1/(2 + d)}
 	\end{equation*}
@@
 \end{enumerate}
-
-We use~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_estimation} to ensure that the approximation error (bias) of $\wh{f}_{\LE}$ is of the correct order.
-On the other hand, bounding the estimation error of $\wh{f}_{\LE}$ -- both in expectation and in probability-- is straightforward.
-Together, bounds on these two sources of error imply Theorem~\ref{thm:laplacian_eigenmaps_estimation1}.
+On the other hand bounding the estimation error of $\wh{f}_{\LE}$ is straightforward. Together, bounds on these two sources of error imply Theorem~\ref{thm:laplacian_eigenmaps_estimation1}.
 \begin{theorem}
 	\label{thm:laplacian_eigenmaps_estimation1}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
-Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_estimation}.
-Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
-	\begin{equation*}
-	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4 + d)} n^{-2/(2 + d)}
-	\end{equation*}
-	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression} with $f_0 \in H^1(\Xset,M)$.
+Suppose that the neighborhood graph $G_{n,r}$ is computed with a radius $r$ which satisfies~\ref{asmp:le_kernel_radius_estimation}, and the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ with $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$.
+Then for all $n$ sufficiently large,
+	\begin{equation*}
+	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{d/(2 + d)} n^{-2/(2 + d)}
+	\end{equation*}
+	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp\{-\kappa\}$.
 \end{theorem}
 
-\begin{itemize}
-	\item Discuss the choice of radius $r$ imposed by~\ref{asmp:kernel_radius_estimation}.
-	\item Comment on the distinction between error rates as measured by the norm in $L^2(P_n)$ vs. $L^2(P)$.
-	\item \textbf{Proof techniques}: to prove results we rely on both (a) classical bias-variance decomposition of orthogonal series estimators (see e.g.
-\textcolor{red}{Tsybakov}), and (b) pointwise and variational results on neighborhood graph Laplacian semi-norms and eigenvalues (see e.g.
-\textcolor{red}{Belkin04, Belkin08, GarciaTrillos18, Calder19})
-\end{itemize}
-
-Theorem~\ref{thm:laplacian_smoothing_estimation1} is analogous to Theorem~\ref{thm:laplacian_eigenmaps_estimation1}, except it deals with the Laplacian smoothing estimator $\wt{f}_{\LS}$ rather than the Laplacian eigenmaps estimator $\wh{f}_{\LE}$.
+In Theorem~\ref{thm:laplacian_smoothing_estimation1} we consider the Laplacian smoothing estimator $\wt{f}_{\LS}$, and draw a similar conclusion to Theorem~\ref{thm:laplacian_eigenmaps_estimation1}.
+One difference: to establish Theorem~\ref{thm:laplacian_smoothing_estimation1}, we will require (much) tighter restrictions on $r$ than were required for Theorem~\ref{thm:laplacian_eigenmaps_estimation1}.
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{1}
+	\item 
+	\label{asmp:ls_kernel_radius_estimation}
+	The sequence of radii $r = r(n)$ satisfies
+	\begin{equation*}
+	\biggl(\frac{\log n}{n}\biggr)^{1/d} \lesssim r(n) \lesssim n^{-3/(4 + 2d)} M^{(d - 4)/(4 + 2d)}
+	\end{equation*}
+	for all $n$ sufficiently large.
+\end{enumerate}
 \begin{theorem}
 	\label{thm:laplacian_smoothing_estimation1}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
-Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_estimation}.
-Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when we set $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$ the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
-	\begin{equation*}
-	\Bigl\|\wh{f}_{\LS} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4 + d)} n^{-2/(2 + d)}
-	\end{equation*}
-	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d < 4$ and $f_0 \in H^1(\Xset,M)$.
+Suppose that that the neighborhood graph $G_{n,r}$ is computed with a radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_estimation},  and the Laplacian smoothing estimator $\wh{f}_{\LS}$ with $\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$ and $s = 1$.
+Then for all $n$ sufficiently large,
+	\begin{equation*}
+	\Bigl\|\wh{f}_{\LS} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{d/(2 + d)} n^{-2/(2 + d)}
+	\end{equation*}
+	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp\{-c M^{d/(2d + 4)} n^{d/(2+d)}\}$.
 \end{theorem}
+To summarize: except on a set of small probability, when appropriately tuned the Laplacian eigenmaps ($\wh{f}_{\LE}$) and Laplacian smoothing ($\wh{f}_{\LS}$) estimators have asymptotic in-sample mean-squared error within a constant factor of the minimax risk.
+Some remarks:
 
 \begin{itemize}
-	\item \textbf{Proof techniques}: Because the solution to~\eqref{eqn:laplacian_smoothing} has an analytic form, we can rely on a bias-variance decomposition---as opposed to empirical process theory---to derive~Theorem~\ref{thm:laplacian_smoothing_estimation1}.
+	\item The range of radii $r$ imposed by~\ref{asmp:le_kernel_radius_estimation} and~\ref{asmp:ls_kernel_radius_estimation} are compatible with practice, where by far the most common choice of radius is the connectivity threshold $r(n) \asymp (\log(n)/n)^{1/d}$, chosen to make the graph $G_{n,r}$ as sparse as possible while still being connected.
+	\item For Laplacian eigenmaps, the choice of $\kappa$ is made to balance the squared-bias and variance terms, i.e. chosen so that with high probability
+	\begin{equation*}
+	\frac{f_0^T \Lap_{n,r} f_0}{n \lambda_{\kappa}(G_{n,r})} \lesssim \frac{\kappa}{n}.
+	\end{equation*}
+	Unlike in the analysis of the Fourier series projection estimator, here the (conditional on $\mathbf{X}$) bias is itself a random variable.
+To upper bound it, we rely on the pointwise and spectral convergence of neighborhood graph Laplacian semi-norms and eigenvalues, as established in Section~\ref{sec:graph_sobolev_classes}.
+	\item A similar set of conclusions hold with respect to the choice of $\rho$ for the Laplacian smoothing estimator $\wt{f}_{\LS}$, where we seek to balance
+	\begin{equation*}
+	\frac{\rho}{n} f_0^T \Lap_{{n,r}} f_0 \lesssim \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k}(G_{n,r}) + 1\bigr)^2}.
+	\end{equation*}
+	The notable differences are: (i) now both the (conditional on $\mathbf{X}$) bias and variance are random, and (ii) to balance bias and variance we require control on all the eigenvalues $\lambda_1(G_{n,r}),\ldots,\lambda_n(G_{n,r})$.
+The second point is the source of the tighter upper bound on $r$ we require in~\ref{asmp:ls_kernel_radius_estimation}.
+	\item Our graph-based estimators are defined only at the design points $(X_1,\ldots,X_n)$ and we therefore measure error using the in-sample mean-squared error $\norm{\cdot}_n^2$.
+Because we work in the random design framework, the in-sample error is a random variable, and our bound is thus a bound in high probability.
+That being said, it is possible to smoothly extend $\wh{f}_{\LE}$ and $\wt{f}_{\LS}$ to be defined on all of $\Xset$, or indeed all of $\Reals^d$---the canonical extension being the \emph{Nystrom extension} \textcolor{red}{(reference)}---and evaluate the error of the extension in $\Leb^2(\Xset)$ norm.
+Assuming that the extension of our estimators and the regression function $f_0$ are suitably smooth, tools from empirical process theory will guarantee that the $\Leb^2(\Xset)$ error is not too much greater than the in-sample error, but we do not further pursue the details here.
 	\item A key difference between Theorems~\ref{thm:laplacian_eigenmaps_estimation1} and~\ref{thm:laplacian_smoothing_estimation1} is that the latter holds only when $d \leq 4$.
-This restriction is unsurprising, and reflects a fundamental drawback in (penalized) least squares estimators more generally.
-The underlying phenomenon is similar to that elucidated by the seminal papers~\cite{birge1993} and~\cite{birge1998}: when the dimension $d$ is large enough, the parameter space is too large, and (penalized) least squares approaches are no longer minimax optimal.
-It is interesting to note that we see a gap emerge between the graph- (Laplacian smoothing) and continuum (tensor product spline) estimator when $d = 3$: the former continues to be minimax optimal, whereas the latter no longer is.
-We should also mention that in a related context (estimation over a $d$-dimensional lattice graph), \cite{sadhanala2016} require the same restriction $d \leq 4$.
+This restriction reflects a fundamental drawback in (penalized) least squares estimators more generally.
+The underlying phenomenon is similar to that elucidated by the seminal papers~\cite{birge1993} and~\cite{birge1998}: when the dimension $d$ is large enough, the parameter space---in their case $C^1(\Xset,M)$, in our case $H^1(G_{n,r},M)$---is too large, and (penalized) least squares approaches are no longer minimax optimal.
+It is interesting to note that we see a gap emerge between the graph- (Laplacian smoothing) and continuum- (smoothing spline) estimators when $2 \leq d \leq 3$: the former continues to be minimax optimal, whereas the latter is not even well-defined.
+We should also mention that in a related context (estimation over a $d$-dimensional lattice graph), \cite{sadhanala2016} arrive at the same restriction $d < 4$.
+As in that work, when $d = 4$ our analysis results in a bound on risk within a $\log n$ factor of the minimax risk (in our context, we set $r \asymp (\log n/n)^{1/4}$ to derive this rates.)
 \end{itemize}
 
-\subsection{Minimax optimal hypothesis testing}
-
-Under appropriate conditions, the Laplacian eigenmaps test statistic $T_{\LE}$ has non-trivial power over all $f_0 \in H^1(P,M)$ that are separated from $0$ in $\Leb^2(P)$ norm by at least the minimax critical radius.
-These conditions include~\ref{asmp:kernel_form} and~\ref{asmp:smooth_density}, but not~\ref{asmp:kernel_radius_estimation}.
-Instead, we will require the following scaling of the neighborhood graph radius.
-\begin{enumerate}[label=(K\arabic*)]
+\subsubsection{Testing}
+
+Let us define two simple tests, the first using $T_{\LE}$ and the second using $T_{\LS}$. For $b \geq 1$, define the thresholds $\wh{t}_b$ and $\wt{t}_b$ to be
+\begin{equation*}
+\wh{t}_{b} := \frac{\kappa}{n} + \frac{2b\sqrt{\kappa}}{n},~~ \wt{t}_{b} := \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k(G_{n,r}) + 1\bigr)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k(G_{n,r}) + 1\bigr)^4}}.
+\end{equation*}
+Then let the Laplacian eigenmaps test $\wh{\phi}_{\LE} := \1\bigl\{T_{\LE} \leq \wh{t}_b\bigr\}$, and the Laplacian smoothing test $\wt{\phi}_{\LS} := \1\bigl\{T_{\LS} \leq \wt{t}_b\bigr\}$; clearly the tests $\wh{\phi}_{\LE}$ and $\wt{\phi}_{\LS}$ depend on $b$, but we suppress this notationally.
+The number $b$---or more accurately $1/b$---reflects the tolerated Type I and Type II error.
+In fact, some standard computations show that the thresholds $\wh{t}_b$ ($\wt{t}_b$) are sufficiently large to control the Type I error of $\wh{\phi}_{\LE}$ ($\wt{\phi}_{\LS}$),
+\begin{equation*}
+\Ebb_0\bigl[\wh{\phi}_{\LE}\bigr],~~\Ebb_0\bigl[\wt{\phi}_{\LS}\bigr] \leq \frac{1}{b^2};
+\end{equation*}
+we relegate these computations to the appendix (see Lemmas~\ref{lem:le_fixed_graph_testing} and~\ref{lem:ls_fixed_graph_testing}).
+It remains to show that these tests have small Type II error whenever $f_0$ is separated from $0$ by more than the critical radius $\epsilon_n^{\star}(\mc{H}^s(\Xset,M))$, which we recall was defined in~\eqref{eqn:critical_radius}.
+We proceed in a parallel manner to Section~\ref{subsubsec:first_order_smoothness_estimation}, starting first with the Laplacian eigenmaps test and moving on to the Laplacian smoothing test.
+
+Just as in the estimation context, to analyze the risk of the Laplacian eigenmaps test we will want to invoke Lemma~\ref{lem:first_order_graph_sobolev_seminorm} and Corollary~\ref{cor:neighborhood_eigenvalue} to upper bound the bias, and so we will need to restrict the scaling of $r(n)$.
+\begin{enumerate}[label=(R\arabic*)]
 	\setcounter{enumi}{2}
 	\item 
-	\label{asmp:kernel_radius_testing}
-	The radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
-	\begin{equation*}
-	n^{-1/d} \leq r(n) \leq n^{-2/(4 + d)}
+	\label{asmp:le_kernel_radius_testing}
+	The sequence of radii $r = r(n)$ satisfies
+	\begin{equation*}
+	\biggl(\frac{\log n}{n}\biggr)^{1/d} \lesssim r(n) \lesssim M^{-4/(4 + d)}n^{-2/(4 + d)}
 	\end{equation*}
@@
 \end{enumerate}
-
+When the sequence of radii $r$ do scale in such a fashion, we establish in Theorem~\ref{thm:laplacian_eigenmaps_testing1} that the Laplacian eigenmaps test achieves, up to constants, the minimax critical radius $n^{-4/(4 + d)}$ for all $f_0 \in H^1(\Xset,M)$.
 \begin{theorem}
 	\label{thm:laplacian_eigenmaps_testing1}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d \leq 3$. Suppose that $f_0 \in H^1(P,M)$, that
-	$P$ satisfies~\ref{asmp:bounded_density}, and that the neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_testing}.
-	Then for the Laplacian eigenmaps test statistic $T_{\LE}$ as in~\eqref{eqn:laplacian_eigenmaps_test} with $\kappa = n^{2d/(4 + d)} M^{4d/(4 + d)}$, the following statement holds true for all $n$ sufficiently large: there exist constants $C_1$ and $C_2$ which do not depend on $f_0$ such that if
-	\begin{equation}
-	\label{eqn:laplacian_eigenmaps_testing}
-	\norm{f_0}_{\Leb^2(P)}^2 \geq \frac{C_1}{\delta^2} M^{2d/(4 + d)} n^{-4/(4 + d)}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d < 4$ and $f_0 \in H^1(\Xset,M)$ for $M \leq n^{(4 - d)/(4 + d)}$.
+Suppose that the neighborhood graph $G_{n,r}$ is computed with a radius $r$ which satisfies~\ref{asmp:le_kernel_radius_testing}, and the Laplacian eigenmaps test $\wh{\phi}_{\LE}$ with $\kappa = n^{2d/(4 + d)} M^{4d/(4 + d)}$ and any $b \geq 1$.
+Then for all $n$ sufficiently large, the following statement holds: if
+	\begin{equation}
+	\label{eqn:laplacian_eigenmaps_testing1}
+	\norm{f_0}_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2d/(4 + d)} n^{-4/(4 + d)}
 	\end{equation}
@@
 	\begin{equation*}
-	\Pbb \biggl(T_{\LE} \geq \frac{\kappa}{n} + \frac{1}{\delta}\sqrt{\frac{2\kappa}{n^2}}\biggr) \geq 1 - C_2\Bigl(\delta + r^{-d} \exp\{- c n r^d\}\Bigr).
+	\Ebb_{f_0}\Bigl[1 - \wh{\phi}_{\LE}\Bigr] \leq C \Bigl(\frac{1}{b} + r^{-d} \exp\{- c n r^d\}\Bigr).
 	\end{equation*}
 \end{theorem}
-\begin{itemize}
-	\item Explain why Theorem~\ref{thm:laplacian_eigenmaps_testing1} implies minimax optimality.
-	\item Include results regarding the low-smoothness regime (i.e. the case when $d \geq 4$.)
-\end{itemize}
-
-\section{Higher-order smoothness classes}
-\label{sec:higher_order_smoothness_classes}
-The estimation and testing rates derived in Section~\ref{sec:minimax_optimal_graph_Laplacian_methods} reflect a bonafide curse of dimensionality: as  the dimension $d$ gets larger, inference quickly becomes (much) harder.
-As covered in Section~\ref{sec:problem_setup_and_background}, it is well known that by making additional structural assumptions---either on the regression function $f_0$, the distribution $P$, or both---one can derive improved minimax rates.
-In the following two sections, we show that under such assumptions, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} can in certain situations obtain the faster (known) minimax rates in the nonparametric estimation and testing problems.
-
-In this section, we will assume that $f_0 \in H^s(P,M)$ for some integer $s > 1$. In other words, we will assume that $f_0$ has at least two smooth derivatives.
-
-\subsection{Minimax optimal estimation}
-\label{subsec:minimax_estimation_higher_order}
-Under the following three conditions, the estimation error of the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves the minimax optimal rate for any $f_0 \in H^s(P,M)$.
-\begin{enumerate}[label=(K\arabic*)]
-	\setcounter{enumi}{2}
+
+Theorem~\ref{thm:laplacian_smoothing_testing1} is analogous to Theorem~\ref{thm:laplacian_eigenmaps_testing1}, except it deals with the Laplacian smoothing test statistic $T_{\LS}$ as opposed to $T_{\LE}$.
+As in the estimation setting, we will require a tight range of scalings for $r$.
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{3}
 	\item 
-	\label{asmp:kernel_radius_higher_order}
-	\textbf{Neighborhood graph radius}:
-	The sequence of radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
-	\begin{equation*}
-	n^{-1/(2(s - 1) + d)}\leq r(n) \leq M^{2/(2s + d)}n^{-1/(2s + d)}
+	\label{asmp:ls_kernel_radius_testing}
+	The sequence of radii $r = r(n)$ satisfies
+	\begin{equation*}
+	\biggl(\frac{\log n}{n}\biggr)^{1/d} \lesssim r(n) \lesssim M^{(16 - d)/(8 + 2d)}n^{(d - 20)/(32 + 8d)}
 	\end{equation*}
@@
 \end{enumerate}
-\begin{enumerate}[label=(F\arabic*)]
-	\setcounter{enumi}{0}
+
+\begin{theorem}
+	\label{thm:laplacian_smoothing_testing1}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d < 4$ and $f_0 \in H^1(\Xset,M)$ for $M \leq n^{(4 - d)/(4 + d)}$.
+Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}, and the Laplacian smoothing test $\wh{\phi}_{\LS}$ with $\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$, $s = 1$, and any $b \geq 1$.
+Then for all $n$ sufficiently large, the following statement holds: if
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_testing1}
+	\norm{f_0}_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2d/(4 + d)} n^{-4/(4 + d)}
+	\end{equation} 
+	then
+	\begin{equation*}
+	\Ebb_{f_0}\Bigl[1 - \wt{\phi}_{\LS} \Bigr] \leq C\Bigl(\frac{1}{b}\bigl[1 + n^{-d(4 + d)}M^{-2d/(4 + d)}\bigr] + \frac{1}{r^d}\exp\bigl\{-cnr^d\bigr\}\Bigr)
+	\end{equation*}
+\end{theorem}
+
+A few remarks:
+\begin{itemize}
+	\item As mentioned in Section~\ref{subsec:sobolev_spaces}, when $d \geq 4$ we do not have minimax testing rates over $H^1(\Xset,M)$; these Sobolev balls include quite unsmooth functions, indeed functions without bounded $4$th moment, and it seems likely that no test---graph-based or otherwise--- can even be consistent against all alternatives in these classes.
+However, in Section~\ref{subsec:minimax_higher_order_smoothness} we derive upper bounds over those functions in $H^1(\Xset,M)$ which \emph{do} have bounded $4$th moment, in fact over all functions $f \in \Leb^4(\Xset)$.
+	\item To compute the threshold $\wt{t}_b$, one must know each of the eigenvalues $\lambda_1(G_{n,r}),\ldots,\lambda_n(G_{n,r})$.
+This is computationally more onerous than computing any of the estimators/statistics~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test}, which involve either finding only the first $\kappa$ eigenvalue/eigenvector pairs (Laplacian eigenmaps) or solving a linear system (Laplacian smoothing).
+That being said, this is not a serious impediment: in practice using asymptotically-optimal thresholds like $\wh{t}_b$ or $\wt{t}_b$ results in substantial loss of efficiency for moderate values of $n$, and we would make the standard recommendation to calibrate via permutation.
+\end{itemize}
+
+We now briefly set the stage for the rest of this Section.
+The estimation and testing rates established in Theorems~\ref{thm:laplacian_eigenmaps_estimation1}-\ref{thm:laplacian_smoothing_testing1} reflect a bonafide curse of dimensionality: as the dimension $d$ gets larger, inference quickly becomes (much) harder.
+As covered in Section~\ref{sec:problem_setup_and_background}, it is well known that by making additional structural assumptions---either on the smoothness of the regression function $f_0$, the dimension of the domain $\Xset$, or both---one can derive improved minimax rates.
+In the following two subsections, we show that under such assumptions, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} and test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} can in certain situations obtain the well-known faster minimax rates in the nonparametric estimation and testing problems.
+We treat these cases separately---from each other, and from the first-order full-dimensional case just covered---because we will require additional regularity conditions for the analysis to go through.
+
+\subsection{Higher-order smoothness}
+\label{subsec:minimax_higher_order_smoothness}
+
+In this section, we will assume that $f_0 \in H^s(\Xset,M)$ for some integer $s > 1$---in other words, that $f_0$ has at least $s$ (partial) weak derivatives bounded in $\Leb^2$ norm---and upper bound the risk of the Laplacian eigenmaps procedures.
+
+\subsubsection{Estimation}
+\label{subsec:minimax_estimation_higher_order}
+Now we would like to invoke Theorem~\ref{thm:graph_sobolev_seminorm}---as opposed to Lemma~\ref{lem:first_order_graph_sobolev_seminorm}---to upper bound the bias of $\wh{f}_{\LE}$: as a result, we will require that $r(n)$ satisfy the conditions of that Theorem.
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{4}
 	\item 
-	\label{asmp:boundary} \textbf{Boundary condition}; The regression function $f_0$ is zero-trace. In other words, $f_0 \in H_0^s(P,M)$. 
+	\label{asmp:kernel_radius_higher_order_estimation}
+	\textbf{Neighborhood graph radius}:
+	The sequence of radii $r = r(n)$ satisfies
+	\begin{equation*}
+	n^{-1/(2(s - 1) + d)} \lesssim r(n) \lesssim M^{-2/(2s + d)}n^{-1/(2s + d)}
+	\end{equation*}
+	for all $n$ sufficiently large.
 \end{enumerate}
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{1}
+The lower bound in~\ref{asmp:kernel_radius_higher_order_estimation} is now meaningfully larger than the connectivity threshold $r \asymp (\log n/n)^{1/d}$.
+However, fundamentally speaking the assumption is still a mild one: we as the user are free to construct the neighborhood graph $G_{n,r}$ however we'd like.
+In Theorem~\ref{thm:laplacian_eigenmaps_estimation_higher_order}, we show that under such a scaling of $r$ and certain regularity conditions, the Laplacian eigenmaps estimator is, up to constants, asymptotically rate-optimal.
+
+\begin{theorem}
+	\label{thm:laplacian_eigenmaps_estimation_higher_order}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $f_0 \in H_0^s(\Xset,M)$ and $p \in C^{s - 1}(\Xset,p_{\max})$.
+Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:kernel_radius_higher_order_estimation}, and the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ with $\kappa = M^{2/(2s + d)}n^{d/(2s + d)}$.
+Then for all $n$ sufficiently large,
+	\begin{equation*}
+	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4s + d)} n^{-2s/(2s + d)}
+	\end{equation*}
+	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp\{-\kappa\}$.
+\end{theorem}
+
+We have already commented on why we require that $p$ belong to $C^{s-1}(\Xset,M)$ and that $f_0$ be zero-trace, in Section~\ref{subsec:higher_order_graph_sobolev_seminorm}.
+However, it is worth pointing out that similar restrictions are also familiar in the broader context of orthogonal series estimators in nonparametric regression.
+For instance, in \textcolor{red}{(Tsybakov, other references)} it is assumed that $f_0 \in H_{\mathrm{per}}^s(\Xset,M)$.
+Additionally, in \textcolor{red}{(Mukherjee16)}, one of the relatively few treatments of orthogonal series regression with random design sampled from an unknown density, it is assumed that $p \in H^s(\Xset,M)$.
+
+\subsubsection{Testing}
+\label{subsec:minimax_testing_higher_order}
+
+Under similar regularity conditions, the Laplacian eigenmaps test statistic $T_{\LE}$ has non-trivial power whenever $f_0 \in H_0^s(\Xset,M)$ is separated from $0$ in $\Leb^2(\Xset)$ norm by at least (a constant times) the minimax critical radius.
+Additionally, we will require the following scaling of the neighborhood graph radii.
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{5}
 	\item 
-	\label{asmp:smooth_density}\textbf{$\beta$-Holder density}: The design distribution $P$ has a density $p \in C^{\beta}([0,1]^d;p_{\max})$ for some $\beta \geq 0$. 
-\end{enumerate}
-
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_estimation}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
-Suppose that $f_0 \in H_0^s(P,M)$, that $P$ satisfies~\ref{asmp:smooth_density} with $\beta = s - 1$, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_higher_order}.
-Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as in~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2s + d)}n^{d/(2s + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
-	\begin{equation*}
-	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4s + d)} n^{-2s/(2s + d)}
-	\end{equation*}
-	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
-\end{theorem}
-
-We note that the assumption~\ref{asmp:kernel_radius_higher_order} is rather mild, since the user can choose how to construct the neighborhood graph $G_{n,r}$.
-The assumptions~\ref{asmp:boundary} and~\ref{asmp:smooth_density} are more stringent.
-That being said, some restrictions on the boundary behavior of the regression function $f_0$ are typical in the analysis of orthogonal series estimators (see e.g.
-\textcolor{red}{Tsybakov}).
-On the other hand, restrictions on the smoothness of the density $p$ are also typical when analyzing orthogonal series estimators under a random design setup (see e.g.
-\textcolor{red}{Mukherjee16}).
-
-\begin{enumerate}
-	\item \textcolor{red}{(TODO)}: Comment a bit on the proof strategy; i.e., 
-\end{enumerate}
-
-\subsection{Minimax optimal hypothesis testing}
-\label{subsec:minimax_testing_higher_order}
-
-Under appropriate conditions, the Laplacian eigenmaps test statistic $T_{\LE}$ has non-trivial power over all $f_0 \in H_0^s(P,M)$ that are separated from $0$ in $\Leb^2(P)$ norm by at least the minimax critical radius.
-These conditions include~\ref{asmp:kernel_form} and~\ref{asmp:smooth_density}.
-Additionally, we will require the following scaling of the neighborhood graph radii.
-\begin{enumerate}[label=(K\arabic*)]
-	\setcounter{enumi}{2}
-	\item 
-	\label{asmp:kernel_radius_testing}
-	The radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
-	\begin{equation*}
-	n^{-1/(2(s - 1) + d)} \leq r(n)~~\textrm{and}~~ r(n) \leq n^{-2/(4s + d)}
+	\label{asmp:kernel_radius_higher_order_testing}
+	The sequence of radii $r = r(n)$ satisfies
+	\begin{equation*}
+	n^{-1/(2(s - 1) + d)} \leq r(n)~~\textrm{and}~~ r(n) \leq M^{-4/(4s + d)}n^{-2/(4s + d)}
 	\end{equation*}
@@
 In one sense this is still a mild condition, in that the neighborhood graph radius is still a tuning parameter to be selected by the user.
-However, some elementary calculations show that when $d > 4$, it is in fact the case that $n^{-1/(2(s - 1) + d)} > n^{-2/(4s + d)}$; hence no choice of $r$ can satisfy~\ref{asmp:kernel_radius_testing}.
+However, some elementary calculations show that when $d > 4$, it is in fact the case that $n^{-1/(2(s - 1) + d)} \gtrsim n^{-2/(4s + d)}$; hence no choice of $r$ can satisfy~\ref{asmp:kernel_radius_higher_order_testing}, unless $M$ is very small.
 As a result, our theory applies only when $d \leq 4$.
@@
 \begin{theorem}
-	\label{thm:laplacian_eigenmaps_testing}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}. Suppose that $d \leq 4$, that $f_0 \in H_0^s(P,M)$, that
-	$P$ satisfies~\ref{asmp:smooth_density} with $\beta = s - 1$, and that the neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_testing}.
-	Then for the Laplacian eigenmaps test statistic $T_{\LE}$ given by~\eqref{eqn:laplacian_eigenmaps_test} with tuning parameter $\kappa = n^{2d/(4s + d)} M^{4d/(4s + d)}$, the following statement holds true for all $n$ sufficiently large: there exist constants $C_1$ and $C_2$ which do not depend on $f_0$ such that if
+	\label{thm:laplacian_eigenmaps_testing_higher_order}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d \leq 4$, $f_0 \in H_0^s(\Xset,M)$ for some $M \leq n^{(4 - d)/(4 + d)}$, and $p \in C^{s - 1}(\Xset,p_{\max})$.
+Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:kernel_radius_higher_order_testing}, and the Laplacian eigenmaps test $\wh{\phi}_{\LE}$ with $\kappa = n^{2d/(4s + d)} M^{4d/(4s + d)}$ and any $b \geq 1$.
+Then for all $n$ sufficiently large, the following statement holds: if
 	\begin{equation}
 	\label{eqn:laplacian_eigenmaps_testing}
-	\norm{f_0}_{\Leb^2(P)}^2 \geq \frac{C_1}{\delta^2} M^{2d/(4s + d)} n^{-4s/(4s + d)}
+	\norm{f_0}_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2d/(4s + d)} n^{-4s/(4s + d)}
 	\end{equation}
@@
 	\begin{equation*}
-	\Pbb \biggl(T_{\LE} \geq \frac{\kappa}{n} + \frac{1}{\delta}\sqrt{\frac{2\kappa}{n^2}}\biggr) \geq 1 - C_2\Bigl(\delta + r^{-d} \exp\{- c n r^d\}\Bigr).
+	\Ebb_{f_0}\bigl[1 - \wh{\phi}_{\LE}\bigr] \leq C \Bigl(\frac{1}{b} + r^{-d} \exp\{- c n r^d\}\Bigr).
 	\end{equation*}
@@
 \begin{itemize}
-	\item \textbf{Test based on the Laplacian smoothing test statistic}. 
-	\item \textbf{Low-smoothness regime}. Include results on minimax testing when $4s < d$.
+	\item \textbf{Low-smoothness regime}. \textcolor{red}{(TODO): Include results on minimax testing when $4s \leq d$.}
 	\item \textbf{Head off potential criticism}.
-Acknowledge frankly that there is a regime---when $4 < d < 4s$---where minimax testing rates are known, but for which we cannot prove that either of our test statistics achieve these rates.
+\textcolor{red}{(TODO): Acknowledge frankly that there is a regime---when $4 < d < 4s$---where minimax testing rates are known, but for which we cannot prove that either of our test statistics achieve these rates.
+Our analysis will give some rates, but they will be suboptimal.}
+	\item \textbf{Whither Laplacian smoothing?} In the higher-order smoothness case, our analysis on the Laplacian smoothing methods is no longer tight enough, for all but a few values of $s$ and $d$, to imply minimax rate-optimality.
+We do not know whether a tighter analysis is possible, or whether the method itself is defective, and leave the matter to future work.
 \end{itemize}
 
-\section{Manifold Assumption}
-\label{sec:manifold_assumption}
+\subsection{Manifold Assumption}
+\label{subsec:manifold_assumption}
+
+\textcolor{red}{(TODO)}
 
 \section{Discussion}
-
+\label{sec:discussion}
+We have therefore shown that the properties of graph Sobolev classes established in Section~\ref{sec:graph_sobolev_classes} imply that graph-based procedures with high probability asymptotically achieve rate-optimal minimax risk, under appropriate regularity conditions.
 \begin{itemize}
+	\item \textbf{Weaker conditions than required for ``continuum'' methods.}  We would like to emphasize one point regarding these regularity conditions: our analysis of graph-based procedures requires weaker assumptions on the design distribution $P$ than is needed to prove the equivalent results for their continuum counterparts.
+We require only that the density $p$ be bounded above and below, and that the domain $\Xset$ be Lipschitz (in the full-dimensional case) or smooth (in the manifold case), but crucially we do not assume that either the domain or density be known.
+This is in contrast with the continuum methods~\eqref{eqn:orthogonal_series} and~\eqref{eqn:smoothing_spline}, and it quantifies in a precise sense an advantage of graph-based methods over their continuum counterparts: the former automatically learn and leverage the shape of $p$ and $\Xset$ in a way the latter do not.
 	\item \textbf{Relationships between $s$ and $d$}.
@@
 Thus we have an accumulation of error problem.
-	\item \textbf{Possible extensions}.
+	\item 
+		 In order to prove optimality of the continuum methods~\ref{eqn:orthogonal_series} and~\ref{eqn:smoothing_spline}, we require much stronger information on the design distribution $P$ than we do for our graph-based procedures.
+	\item \textbf{Possible extensions}. Methodologically: E.g. kNN graphs, different non-linear embedding schemes, etc. Theoretically: relaxing~\ref{asmp:bounded_density}, relaxing boundary conditions.
+	\item \textbf{Boundary conditions}.
+Refer to \textcolor{red}{Belkin12} for a more precise description of the boundary conditions imposed by Laplacian eigenmaps, and \textcolor{red}{Ting18} for a catalog of potential solutions.
+Make clear that to ``fix'' the boundary bias problem---i.e.
+to obtain sharp rates for functions $f$ which do not satisfy $f \in H_0(\Xset)$, or at the least $f$ satisfying some Neumann boundary condition---likely requires a change of estimators, and is beyond the scope of this work.
 \end{itemize}
@@
 
-\clearpage
-To prove Theorems~\ref{thm:laplacian_eigenmaps_estimation},\textcolor{red}{...}, we would like to apply Lemmas~\ref{lem:fixed_graph_estimation},\textcolor{red}{...} in the random design context, i.e.
-to the random geometric graph $G_{n,r}$.
-To do so, we must establish the following (high probability) bounds: (i) an upper bound on the graph Sobolev seminorm of the regression function $f_0 \Lap_{n,r}^s f_0$, (ii) a lower bound on the Laplacian eigenvalue $\lambda_{\kappa}(G_{n,r})$ as a function of $\kappa$, and (iii) a lower bound on the empirical norm $\norm{f}_{n}$ as a function of $\norm{f}_{L^2(P)}$.
-
-\section{Bounds on the graph Sobolev semi-norm}
+\section{Bound on the neighborhood graph Sobolev semi-norm}
 \label{sec:graph_sobolev_seminorm}
-We first bound $f_0^T \Lap_{n,r}^s f_0$ in expectation.
+
+\section{Bounds on neighborhood graph eigenvalues}
+\label{sec:graph_eigenvalues}
+
+\section{Bounds on the empirical norm}
+\label{sec:empirical_norm}
+
+We use Lemma~\ref{lem:empirical_norm_sobolev} to lower bound $\norm{f_0}_n^2$ by (a constant times) the $\Leb^2$ norm of $f$.
+
 \begin{lemma}
-	\label{lem:roughness_functional_expectation_sobolev}
-	Let $\Xset$ be a Lipschitz domain and let $s \geq 1$ be an integer.
-Suppose that \textcolor{red}{$f \in H_0^{s}(P,M)$}, and that $P$ admits a density $p \in C^{s-1}(\Xset,p_{\max})$ for some constant $p_{\max} < \infty$.
-Let $\{r(n)\}$ be a sequence of connectivity radii such that $n^{-1/(2(s - 1) + d)} \leq r(n)$ for all $n$ sufficiently large, and $r(n) \to 0$ as $n \to \infty$.
-Then \textcolor{red}{for any $2$nd-order kernel $K$}, there exists a constant $c$ depending only on $p_{\max}, K_{\max}$, $s$, $d$ and $\Xset$ such that the resulting neighborhood graph Laplacian satisfies
-	\begin{equation}
-	\label{eqn:roughness_functional_expectation_sobolev}
-	\Ebb\bigl[f^T L_{n,r}^s f\bigr] \leq c \cdot n \cdot (nr^{d+2})^s \cdot M^2
-	\end{equation}
-	for all $n$ sufficiently large.
+	\label{lem:empirical_norm_sobolev}
+	Suppose $P$ satisfies~\ref{asmp:bounded_density}.
+Then for any $\delta \leq 1$ and $f \in H^s(\Xset,M)$, there exists a constant $C_6$ which depends only on $s$, $\Xset$, $d$, $p_{\min}$ and $p_{\max}$, such that if
+	\begin{equation}
+	\label{eqn:empirical_norm_sobolev_1}
+	\norm{f}_{\Leb^2(\Xset)} \geq 
+	\begin{cases*}
+	\frac{C_6 M}{\delta} \cdot \max\Bigl\{n^{-1/2},n^{-s/d}\Bigr\},& ~~\textrm{if $2s \neq d$} \\
+	\frac{C_6 M}{\delta} \cdot n^{-a/2},& ~~\textrm{if $2s = d$, for any $0 < a < 1$}
+	\end{cases*}
+	\end{equation}
+	then,
+	\begin{equation}
+	\label{eqn:empirical_norm_sobolev}
+	\norm{f}_n^2 \geq \delta \Ebb\Bigl[\norm{f}_n^2\Bigr] 
+	\end{equation}
+	with probability at least $1 - 5 \delta$.
 \end{lemma}
-
-\section{Bounds on graph eigenvalues}
-\label{sec:graph_eigenvalues}
-Lemma~\ref{lem:neighborhood_eigenvalue} establishes a scaling rate which holds for all eigenvalues $\lambda_{k}(G_{n,r})$, assuming appropriate regularity conditions on $P$.
-\begin{lemma}
-	\label{lem:neighborhood_eigenvalue}
-	Suppose the distribution $P$ satisfies~\ref{asmp:bounded_density}.
-Then there exists constants $c$ and $C$ which depend only on $p_{\min}, p_{\max}$ and $d$, such that with probability at least $1 - C r^{-d} \exp\bigl\{- c n r^d\bigr\}$, for all $2 \leq k \leq n$
-	\begin{equation}
-	\label{eqn:neighborhood_eigenvalue}
-	\lambda_k(G_{n,r}) \geq c \cdot \min\Bigl\{nr^{d + 2}k^{2/d}, nr^d\Bigr\}.
-	\end{equation}	
-\end{lemma}
-As those readers familiar with Sobolev spaces will be aware of, the scaling $k^{2/d}$ is the \textcolor{red}{typical scaling of a Sobolev ellipsoid}, and is thus the scaling of the eigenvalues we desire.
-On the other hand, the second term on the right hand side of~\eqref{eqn:neighborhood_eigenvalue} is an artifact of the well known inequality $\lambda_n(G) \leq 2 \max_{i = 1,\ldots,n} \mathbf{D}_{ii}$--- in words, that the maximum eigenvalue of the Laplacian can never be more than twice the maximum degree of the graph.
-Doing some algebra, we may deduce the conditions on $k$ and $r$ under which the graph eigenvalue follows the desired scaling, i.e.
-$\lambda_k(G_{n,r}) \sim n r^{d + 2}k^{2/d}$.
-
-First, when $k = M^{2/(2s + d)} n^{d/(2s + d)}$, it holds that for any $r \leq M^{-2/(d(2s +d))} n^{-1/(2s + d)}$, 
-\begin{equation}
-\label{eqn:neighborhood_eigenvalue_1}
-\lambda_{k}(G_{n,r}) \geq c \cdot nr^{d+2}{k}^{2/d}
-\end{equation}
-with probability at least $1 - C r^{-d} \exp\{- c n r^d\}$. Second, when $k$
 
@@
 \label{sec:proofs_theorems}
-With the results of Sections~\ref{sec:fixed_graph_error_bounds}-\ref{sec:graph_eigenvalues} established, the proofs of our main theorems are quite straightforward.
-
-\subsection{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation}}
-\label{subsec:laplacian_eigenmaps_estimation_pf}
-
-\paragraph{\eqref{eqn:fixed_graph_estimation_sobolev_rate_2} holds with high probability.}
-
-Since $\Lap_{n,r}$ is a positive semi-definite matrix, $f_0^T \Lap_{{n,r}}^s f_0$ must be non-negative. We may therefore apply Markov's inequality and conclude
-\begin{equation}
-\label{eqn:roughness_functional_probability_sobolev}
-f_0^T \Lap_{{n,r}}^s f_0 \leq  \frac{1}{\delta} \cdot c \cdot n \cdot (nr^{d+2})^s \cdot M^2
-\end{equation}
-with probability at least $1 - \delta$.
-
-\paragraph{\eqref{eqn:fixed_graph_estimation_sobolev_rate_1} holds with high probability.}
-
-\paragraph{Putting the pieces together.}
-
-Let $n$ be sufficiently large so that $n^{-1/(2(s - 1)+ d)} \leq r(n) \leq M^{-2/(d(2s + d))} n^{-1/(2s + d)}$.
-Then by~\eqref{eqn:roughness_functional_probability_sobolev} and~\eqref{eqn:neighborhood_eigenvalue_1}, there exists a set $\mathcal{E}(f_0) \subseteq \Xset^n$ of probability at least $1 - \delta - Cr^{-d}\exp\{-cnr^d\}$ such that whenever ${\bf X} \in \mathcal{E}(f_0)$,
-\begin{equation*}
-f_0^T \Lap_{n,r}^s f_0 \leq C \cdot n \cdot C_n^s \cdot M^2
-\end{equation*}
-and
-\begin{equation*}
-\lambda_{\kappa}(G_{n,r}) \geq c \cdot nr^{d+2}\kappa^{2/d}
-\end{equation*}
-Theorem~\ref{thm:laplacian_eigenmaps_estimation} then follows by Corollary~\ref{cor:fixed_graph_estimation_sobolev_rate}.
+With the results of Sections~\ref{sec:graph_sobolev_classes} and Section~\ref{sec:fixed_graph_error_bounds} in hand, the proofs of our main theorems are quite straightforward.
+In each case, we will proceed by conditioning on the random design, and showing that there exists a high-probability set $\mc{E} \subseteq \Xset^n$ over which the conditional risk is uniformly upper bounded.
+
+\subsection{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation1}}
+\label{subsec:laplacian_eigenmaps_estimation1_pf}
+
+\textcolor{red}{(TODO)}
+
+\subsection{Proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}}
+\label{subsec:laplacian_smoothing_estimation1_pf}
+Let $\mc{E} = \mc{E}(f_0, K,r,\rho,c,C,\delta) \subset \mc{X}^n$ be the set of $\mathbf{X} = (X_1,\ldots,X_n)$ such that
+\begin{enumerate}[(i)]
+	\item
+	\label{pf:laplacian_smoothing_estimation1_0}
+	 $f_0^T \Lap_{n,r}^s f_0 \leq \frac{C}{\delta} n^2 r^{d + 2}M^2$
+	\item 
+	\label{pf:laplacian_smoothing_estimation1_1}
+	$c \cdot \min\{k^{2/d}nr^{d+2},nr^d\} \leq \lambda_k(G_{n,r}) \leq Ck^{2/d}nr^{d + 2} $ for all $2 \leq k \leq n$.
+\end{enumerate}
+We already know, from Lemmas~\ref{lem:first_order_graph_sobolev_seminorm} and~\ref{lem:neighborhood_eigenvalue}, that there exist sufficiently small/large constants $c$ and $C$ such that
+\begin{equation}
+\label{pf:laplacian_smoothing_estimation1_1.5}
+\Pbb(\mc{E}) \geq 1 - \delta - \frac{C}{r^d}\exp\bigl\{-cnr^d\bigr\}
+\end{equation}
+We will now show that for any such $\mathbf{X} \in \mc{E}$, the empirical mean square error is small with high probability.
+Having conditioned on $\mathbf{X}$, we can control the remaining randomness arising from the noise $Y$ by Lemma~\ref{lem:ls_fixed_graph_estimation}, which in this context implies
+\begin{equation}
+\label{pf:laplacian_smoothing_estimation1_2}
+\begin{aligned}
+\Pbb\Biggl(& \frac{1}{n}\bigl\|\wt{f}_{\LS}(G_{n,r}) - f_0\bigr\|_2^2 \leq \frac{2\rho}{n} \bigl(f_0^T \Lap_{n,r} f_0\bigr) + \frac{10}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k}(G_{n,r}) + 1\bigr)^2} \Bigg| \mathbf{X}\Biggr) \\
+& \geq 1 - \exp\Biggl\{-\sum_{k = 1}^{n}\frac{1}{\bigl(\rho \lambda_k(G_{n,r}) + 1\bigr)^2}\Biggr\} \\
+& \geq 1 - \exp\Biggl\{-\sum_{k = 1}^{n}\frac{1}{\bigl(CM^{-4/(2+d)}k^{2/d}n^{-2/(2+d)} + 1\bigr)^2}\Biggr\} \\
+& \geq 1 - \exp\Bigl\{-c M^{d/(2d + 4)} n^{d/(2+d)}\Bigr\}
+\end{aligned}
+\end{equation}
+where the second inequality follows from the lower bound in~\ref{pf:laplacian_smoothing_estimation1_1} and the choice of regularization parameter $\rho = M^{-4/(2+d)}(nr^{d+2})^{-1}n^{-2/(2+d)}$.
+
+We have therefore shown that with high probability, the empirical mean squared error is upper bounded by squared bias and variance terms.
+It remains only to upper bound the squared bias and variance terms.
+A sufficient upper bound on the bias term comes immediately from the upper bound~\ref{pf:laplacian_smoothing_estimation1_0} and our choice of $\rho$,
+\begin{equation}
+\label{pf:laplacian_smoothing_estimation1_2.5}
+\frac{\rho}{n} \bigl(f_0^T \Lap_{{n,r}} f_0\bigr) \leq \frac{C}{\delta} M^{2/(2+d)} n^{-2/(2 + d)}.
+\end{equation}
+
+To upper bound the variance term, we replace the eigenvalues $\lambda_k(G_{n,r})$ by their lower bounds in~\ref{pf:laplacian_smoothing_estimation1_1}, plug in our choice of $\rho$, apply the upper bound $r \leq C n^{-3/(4 + 2d)} M^{(d - 4)/(4 + 2d)}$ given by~\ref{asmp:ls_kernel_radius}, and obtain
+\begin{align}
+\frac{C}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k}(G_{n,r}) + 1\bigr)^2} & \leq \frac{C}{n}\sum_{k = 1}^{n} \biggl\{\frac{1}{(\rho k^{2/d} n r^{d + 2} + 1)^2} + \frac{1}{(\rho n r^{d} + 1)^2} \biggr\} \nonumber \\
+& \leq \frac{C}{n}\sum_{k = 1}^{n} \biggl\{\frac{1}{\bigl(M^{-4/(2 + d)}n^{-2/(2+d)}k^{2/d} + 1\bigr)^2}\biggr\} + Cn^{4/(2 + d)}M^{8/(2+d)}r^{4} \nonumber \\
+& \leq CM^{2d/(2 + d)} n^{-2/(2 + d)} + CM^{8/(2 + d)}n^{(2 - d)/(2 + d)} \sum_{k = k_{\star}}^{n} \Bigl\{\frac{1}{k^{4/d}}\Bigr\}  + CM^{8/(2 + d)}n^{4/(2 + d)}r^4 \nonumber \\
+& \leq  CM^{2d/(2 + d)} n^{-2/(2 + d)} + CM^{8/(2 + d)}n^{(2 - d)/(2 + d)} \sum_{k = k_{\star}}^{n} \Bigl\{\frac{1}{k^{4/d}}\Bigr\}  + CM^{2d/(2 + d)}n^{-2/(2 + d)} \label{pf:laplacian_smoothing_estimation1_3}
+\end{align}
+where $k_{\star} = M^{2d/(2 + d)}n^{d/(2 + d)}$.
+Treating the last term in the above line as a Riemann sum evaluated at the right end points of $[k_{\star},k_{\star} + 1], [n - 1,n]$, we use the fact that such a Riemann sum upper bounds the integral of a montonically nonincreasing function to conclude that
+\begin{align*}
+\sum_{k = k_{\star}}^{n} \frac{1}{k^{4/d}} & \leq \int_{k_{\star}}^{n} \frac{1}{x^{4/d}} \,dx \\
+& = - \frac{1}{(4/d - 1)} x^{-(4/d - 1)} \Big|_{k_{\star}}^{n} \\
+& \leq \frac{1}{(4/d - 1)} k_{\star}^{-(4/d - 1)}
+\end{align*}
+where the equality in the above computation holds because $d < 4$.
+Recalling the definition of $k_{\star}$, plugging this back into~\eqref{pf:laplacian_smoothing_estimation1_3} implies a sufficient upper bound on the variance term.
+Along with~\eqref{pf:laplacian_smoothing_estimation1_1.5}-\eqref{pf:laplacian_smoothing_estimation1_2.5}, this establishes the claim.
+
+\subsection{Proof of Theorem~\ref{thm:laplacian_eigenmaps_testing1}}
+In this proof---as well as in the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing1}---we will need to invoke Lemma~\ref{lem:empirical_norm_sobolev}, and it is in order to satisfy the conditions of this Lemma that we require $M \leq n^{(4 - d)/(4 + d)}$.
+More specifically, by~\eqref{eqn:laplacian_eigenmaps_testing1} along with the restriction $M \leq n^{(4 - d)/(4 + d)}$,
+\begin{equation*}
+\norm{f}_{\Leb^2(\Xset)} \geq \frac{C}{\delta^2} M^{2d/(4 + d)} n^{-4/(4 + d)} \geq \frac{C}{\delta^2} M^2 n^{-2/d}
+\end{equation*}
+and~\eqref{eqn:empirical_norm_sobolev_1} is therefore satisfied when we choose $C \geq C_6$ in~\eqref{eqn:laplacian_eigenmaps_testing1}.
+
+We proceed to prove Theorem~\ref{thm:laplacian_eigenmaps_testing1}.
+Throughout this proof we set $\delta = 1/b$.
+Let $\mc{E}(f_0)$ be the set of $\mathbf{X} \in \Xset^n$ such that~\eqref{eqn:first_order_graph_sobolev_seminorm},~\eqref{eqn:neighborhood_eigenvalue}, and~\eqref{eqn:empirical_norm_sobolev_1} hold: by Lemmas~\ref{lem:first_order_graph_sobolev_seminorm},\ref{lem:neighborhood_eigenvalue} and~\ref{lem:empirical_norm_sobolev}, $\Pbb(\mc{E}(f_0)) \geq 1 - 6\delta - C_4/r^d\exp\{-c_4nr^d\}$.
+For any $\mathbf{X} \in \mc{E}(f_0)$,
+\begin{align*}
+\frac{f_0^T \Lap_{{n,r}} f_0}{n \lambda_k(G_{n,r})} + \frac{4\sqrt{\kappa}}{\delta n} & \leq \frac{(C_1/c_4 + 4)}{\delta}M^{2d/(4 + d)}n^{-4/(4 + d)} \\
+& \leq \frac{\delta}{p_{\min}} \norm{f}_{\Leb^2(\Xset)}^2 \leq \delta \Ebb\bigl[\norm{f}_n^2\bigr] \\
+& \leq \norm{f}_n^2
+\end{align*}
+where the second inequality follows upon choosing $C \geq p_{\min}^{-1}(C_1/c_4 + 4)$ in~\eqref{eqn:laplacian_eigenmaps_testing1}. The claim follows by Lemma~\ref{lem:le_fixed_graph_testing}.
+
+\subsection{Proof of Theorem~\ref{thm:laplacian_smoothing_testing1}}
+Let $\delta$ and $\mc{E}(f_0)$ be the same as in the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing1}.
+For any $\mathbf{X} \in \mc{E}(f_0)$, the conditional variance of $T_{\LS}$ can be upper bounded using~\eqref{eqn:neighborhood_eigenvalue} and~\ref{asmp:ls_kernel_radius_testing},
+\begin{align*}
+\frac{1}{n} \biggl(\sum_{k = 1}^{n} \bigl[\lambda_k(G_{n,r}) + 1\bigr]^{-4}\biggr)^{1/2} & \leq \frac{1}{c_5^2 n} \biggl(M^{\frac{32}{8 + d}}r^8n^{\frac{20+d}{4+d}} + \sum_{k = 1}^{n} \bigl[n^{-\frac{4}{4 + d}}k^{\frac{2}{d}}M^{-\frac{8}{4 + d}} + 1\bigr]^{-4}\biggr)^{1/2} \\
+& \leq \frac{1}{c_5^2 n} \biggl(M^{\frac{4d}{4 + d}}n^{\frac{2d}{4 + d}} + \sum_{k = 1}^{n} \bigl[n^{-\frac{4}{4 + d}}k^{\frac{2}{d}}M^{-\frac{8}{4 + d}} + 1\bigr]^{-4}\biggr)^{1/2} \\
+& \leq \frac{1}{c_5^2 n} \biggl(3 M^{\frac{4d}{4 + d}}n^{2d/(4 + d)}\biggr)^{1/2} \\
+& = \frac{\sqrt{3}}{c_5^2} M^{\frac{2d}{4 + d}} n^{-\frac{4}{4 + d}}.
+\end{align*}
+The last inequality in the above display follows from treating the sum as a Riemann sum over a monotonically decreasing function, which can be upper bounded by an integral as in \ref{subsec:laplacian_smoothing_estimation1_pf}, i.e by setting $k_{\star} = n^{2d/(4+d)}M^{4d/(4 + d)}$ and deriving
+\begin{equation}
+\begin{aligned}
+\label{pf:laplacian_smooting_testing1}
+\sum_{k = 1}^{n} \bigl[n^{-\frac{4}{4 + d}}k^{\frac{2}{d}}M^{-\frac{8}{4 + d}} + 1\bigr]^{-4} & \leq k_{\star} + n^{\frac{16}{4 + d}} M^{\frac{32}{4 + d}}\sum_{k = k_{\star}}^{n} k^{-\frac{8}{d}}\\
+& \leq k_{\star} + n^{\frac{16}{4 + d}} M^{\frac{32}{4 + d}} \int_{k_{\star}}^{n} x^{-\frac{8}{d}} \,dx \\
+& \leq  k_{\star} + \frac{1}{(8/d - 1)}n^{\frac{16}{4 + d}} M^{\frac{32}{4 + d}} k_{\star}^{1 - 8/d} \\
+& \leq 2k_{\star}.
+\end{aligned}
+\end{equation}
+As a result, 
+\begin{align*}
+\frac{2\rho}{n} \bigl(f_0^T \Lap_{n,r} f_0) + \frac{4}{\delta n}\biggl(\sum_{k = 1}^{n} \bigl[\lambda_k(G_{n,r}) + 1\bigr]^{-4}\biggr)^{1/2} & \leq \frac{1}{\delta}\biggl(2C_1 + \frac{4\sqrt{3}}{c_5^2}\biggr)n^{-\frac{4}{4 + d}} M^{\frac{2d}{4 + d}} \\
+& \leq \delta p_{\min} \norm{f}_{\Leb^2(\Xset)}^2 \leq \delta \Ebb\bigl[\norm{f}_n^2\bigr] \\
+& \leq \norm{f}_n^2,
+\end{align*}
+where the second inequality follows upon choosing $C \geq p_{\min}^{-1}(2C_1 + 4\sqrt{3}/c_5^2)$ in~\eqref{eqn:laplacian_smoothing_testing1}.
+We may therefore bound the Type II error of $\wt{\phi}_{\LS}$ by appealing to Lemma~\ref{lem:ls_fixed_graph_testing},
+\begin{align*}
+\Ebb_{f_0}\Bigl[1 - \wt{\phi}_{\LS}\Bigr] & \leq \Ebb_{f_0}\Bigl[1 - \wt{\phi}_{\LS}\big| \mathbf{X} \in \mc{E}(f_0)\Bigr] + 1 - \Pbb\Bigl(\mc{E}(f_0)\Bigr) \\
+& \leq 4 \delta^2 + 8 \delta \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G_{n,r})^s + 1)^4} \Biggr)^{-1/2} + 6\delta + \frac{C_4}{r^d}\exp\{-c_4nr^d\} \\
+& \leq 4 \delta^2 + 16\delta n^{-d(4 + d)}M^{-2d/(4 + d)}  + 6\delta + \frac{C_4}{r^d}\exp\bigl\{-c_4nr^d\bigr\} 
+\end{align*}
+where the last inequality follows from similar reasoning to~\eqref{pf:laplacian_smooting_testing1}, with the inequalities reversed and a factor of $1/2$ instead of $2$ in front.
 

### 2020-09-16 notes/graph_regression2.tex
Completed an edit of the first draph of our graph regression paper; time to merge.

@@
 \end{equation}
-with independent Gaussian noise $\varepsilon_i \sim N(0,1)$.
-Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether or not $f_0 = 0$, i.e whether or not there is a signal present.
-
-In graph based nonparametric regression, the idea is to perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$.
-For a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$.
-The weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ is an $n \times n$ matrix with entries
+with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise.
+Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present.
+
+In graph-based nonparametric regression, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$.
+The neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$.
+The $n \times n$ weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $\mathbf{W}_{ij}$ are given by
 \begin{equation*}
 \label{eqn:neighborhood_graph}
-{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr),
-\end{equation*}
-and the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$.
-Our star operator, the neighborhood graph Laplacian, can then be written as
+{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr).
+\end{equation*}
+Then the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$, and our star operator, the neighborhood graph Laplacian, can be written as
 \begin{equation}
@@
 \label{eqn:laplacian_eigenmaps}
-\wh{f}_{\textrm{LE}} = \sum_{k = 1}^{\kappa} \Dotp{Y}{v_k(G_{n,r})}_n v_{k}(G_{n,r})
-\end{equation}
-where $\kappa \in \{1,...,n\}$ is a tuning parameter.
-(Here $\dotp{u}{v}_n = \frac{1}{n}\sum_{i = 1}^{n} u_i v_i$ for vectors $u,v \in \Reals^n$.) The Laplacian smoothing estimator $\wt{f}_{\LS}$ \citep{smola2003} is a penalized least squares estimator, given by
+\wh{f}_{\textrm{LE}} = \sum_{k = 1}^{\kappa} \biggl\{\Bigl(\frac{1}{n}\sum_{i = 1}^{n} Y_i v_{k,i}(G_{n,r})\Bigr) v_{k}(G_{n,r})\biggr\}
+\end{equation}
+where $\kappa \in \{1,...,n\}$ is a tuning parameter. The Laplacian smoothing estimator $\wt{f}_{\LS}$ \citep{smola2003} is a penalized least squares estimator, given by
 \begin{equation}
 \label{eqn:laplacian_smoothing}
-\wt{f}_{\LS}= \min_{f \in \Reals^n} \Bigl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_{n,r}^s f \Bigr\}
-\end{equation}
-where $\rho > 0$ acts a tuning parameter $f^T \Lap_{{n,r}}^s f$.
-Then, assuming~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are reasonable estimators of $f_0$, the squared empirical norms
-\begin{equation}
-T_{\LE} = \Bigl\|\wh{f}_{\LE}\Bigr\|_n^2 \label{eqn:laplacian_eigenmaps_test}
+\wt{f}_{\LS}= \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_{n,r}^s f \biggr\}
+\end{equation}
+where $\rho > 0$ acts a tuning parameter on the penalty $f^T \Lap_{{n,r}}^s f$.
+Assuming~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are reasonable estimators of $f_0$, the squared empirical norms
+\begin{equation}
+T_{\LE} = \frac{1}{n}\Bigl\|\wh{f}_{\LE}\Bigr\|_2^2 \label{eqn:laplacian_eigenmaps_test}
 \end{equation}
@@
 \begin{equation}
-T_{\LS} = \Bigl\|\wt{f}_{\LS}\Bigr\|_n^2 \label{eqn:laplacian_smoothing_test}
+T_{\LS} = \frac{1}{n}\Bigl\|\wt{f}_{\LS}\Bigr\|_2^2 \label{eqn:laplacian_smoothing_test}
 \end{equation}
@@
 
-There already exist a vast number of methods for both the nonparametric testing and, especially, the nonparametric estimation problems.
-In fact, as we will see in Section~\ref{sec:problem_setup_and_background} some of these methods are actually quite analogous to the graph Laplacian estimation and testing methods we have just introduced.
-Nevertheless, methods such as~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} have a deserved place in the nonparametric regression toolbox.
-Graph-based methods such as these have many advantages: they are easy to implement, fast and stable to compute, generalizable to non-standard data---for instance, text or images, or really any data modality on which it is possible to define a kernel--- and non-standard learning paradigms---such as the semi-supervised setting---and last but not least, have enjoyed widespread practical success.
-
-For these reasons, a substantial body of work has arisen assessing the statistical properties of graph Laplacian based methods.
-Taken as a whole, this work views the graph Laplacian matrix as an estimator of a weighted continuum Laplace operator $\Delta_P$ (see equations~\eqref{eqn:laplace_operator} and~\eqref{eqn:laplace_operator_weighted} in Section~\ref{sec:problem_setup_and_background}), and establishes that it is in fact a \emph{consistent} estimator of $\Delta$, in various respects.
-For instance,  \citet{belkin03,belkin05} show that when $\Pbb$ is the uniform distribution over $\Xset$, $\Lap_{n,r}f \to \Delta f$ for sufficiently smooth functions $f$, \citet{lafon04,hein05} extend the statement to hold when $\Pbb$ is non-uniform, \citet{singer06} sharpens the convergence rates, and \cite{gine06} establish a CLT.
-On the other hand, \citet{belkin07} prove that when $\Pbb$ is the uniform distribution over $\mathcal{X}$, the eigenvalues and eigenvectors of $\Lap_{n,r}$ converge to those of $\Delta$ when $r_n \to 0$ as $n \to \infty$, \citet{garciatrillos18} extend this result to hold for general $\Pbb$, and \cite{calder2019} obtain sharper rates at which $r_n$ may converge to $0$.
+Graph-based methods have a deserved place in the nonparametric regression toolbox.
+Of course, there exist a considerable range of methods for both the nonparametric testing and, especially, the nonparametric estimation problems.
+In fact, as we will see in Section~\ref{sec:problem_setup_and_background} some of these more classical methods are actually quite analogous to~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing}.
+Nevertheless, graph-based methods have many strengths: they are easy to implement; fast and stable to compute; generalizable to non-standard data---for instance, text or images, or really any data modality on which it is possible to define a kernel--- and non-standard learning paradigms---such as the semi-supervised setting.
+Last but not least, they have enjoyed widespread practical success.
+
+For these reasons, a substantial body of work has arisen assessing the statistical properties of graph-based methods.
+Taken as a whole, this work views the graph Laplacian matrix as an estimator of a weighted continuum Laplace operator $\Delta_P$ (see equations~\eqref{eqn:laplace_operator} and~\eqref{eqn:laplace_operator_weighted} in Section~\ref{sec:problem_setup_and_background}), and establishes that when $n \to \infty$ and $r(n) \to 0$, it is in fact a \emph{consistent} estimator of $\Delta_P$.
+This notion of consistency is meant in a few different respects.
+For instance,  \citet{belkin03,belkin05} show that when $P$ is the uniform distribution over $\Xset$, $\Lap_{n,r}f \to \Delta f$ for sufficiently smooth functions $f$, \citet{lafon04,hein05} extend the statement to hold when $P$ is non-uniform, \citet{singer06} sharpens the convergence rates, and \cite{gine06} prove a central limit theorem.
+In another line of work, \citet{belkin07} prove that when $P$ is the uniform distribution over $\mathcal{X}$, the eigenvalues and eigenvectors of $\Lap_{n,r}$ converge to those of $\Delta_P$, \citet{garciatrillos18} extend this result to hold for when $P$, and \cite{calder2019} sharpen the rates at which $r(n)$ may converge to $0$.
 
@@
 In this paper, we confirm that they are.
-We show that when the regression function $f_0$ is suitably smooth---by which we mean that it belongs to a first-order Sobolev ball; roughly speaking, that it possesses weak derivative $Df$ bounded in $L^2(\Xset)$ norm---both~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are with high probability asymptotically rate-optimal estimators, and that both~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} are asymptotically rate-optimal test statistics.
-We also show that under appropriate conditions, these estimators and test statistics can leverage additional assumed structure, either on the regression function $f_0$ or distribution $P$.
-Our fundamental message is a reassuring one: the practitioner who wishes to use neighborhood graph procedures, for any of the various reasons outline above, may do so without fear that they are sacrificing statistical optimality in the process.
-
-\begin{itemize}
-	\item \textbf{Motivate graph Laplacians.}
-
-	\item \textbf{Literature review.} 
-	\item \textbf{Organization.}
-\end{itemize}
+We establish that when the regression function $f_0$ is suitably smooth---by which we mean that it belongs to a first-order Sobolev ball; roughly speaking, that it possesses weak derivative $Df$ bounded in $L^2(\Xset)$ norm---both~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are with high probability asymptotically rate-optimal estimators, and that both~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} are asymptotically rate-optimal test statistics.
+We also show that these estimators and test statistics can leverage additional assumed structure, either on the regression function $f_0$ or the distribution $P$.
+Our fundamental message is a reassuring one: the practitioner who wishes to use neighborhood graph procedures, for any of the various reasons outlined above, may do so without fear that they are sacrificing statistical optimality in the process.
+
+\textcolor{red}{(TODO 1)}: Outline organization of paper here.
 
@@
 \begin{itemize}
-	\item We write $\mathbb{Z}^d = \mathbb{Z} \otimes \cdots \otimes \mathbb{Z}$ (the tensor product is taken a total of $d$ times); the sequence space $\ell^2(\mathbb{Z}^d)$ contains those vectors $\theta: \mathbb{Z}^d \to \Reals$ for which $\sum_{k \in \mathbb{Z}^d} \theta_k^2 < \infty$.
-	\item We denote by $\Leb^2(\Xset)$ the set of square integrable functions $f:\Xset \to \Reals$. \textcolor{red}{Holder}.
-	\item For functions $f,g$ which map from $\Xset$ to $\Reals$, we write $\dotp{f}{g}_n := \frac{1}{n}\sum_{i = 1}^{n} f(X_i) g(X_i)$, which we refer to as the \emph{empirical inner product}, and $\norm{f}_n^2 = \dotp{f}{f}_n$, which we refer to as the (squared) \emph{empirical norm}.
+	\item We write $\mathbb{Z}^d = \mathbb{Z} \otimes \cdots \otimes \mathbb{Z}$---with the tensor product being taken a total of $d$ times--- and likewise $\mathbb{N}^d = \mathbb{N} \otimes \cdots \otimes \mathbb{N}$.
+The sequence space $\ell^2(\mathbb{Z}^d)$ then contains those vectors $\theta: \mathbb{Z}^d \to \Reals$ for which $\sum_{k \in \mathbb{Z}^d} \theta_k^2 < \infty$.
+	\item For $1 \leq p < \infty$, we denote by $\Leb^p(\Xset)$ the set of functions $f:\Xset \to \Reals$ with bounded $p$th moment $\int_{\Xset} \bigl(f(x)\bigr)^p \,dx < \infty$.
+When $p = \infty$, we let $\Leb^{\infty}(\Xset)$ contain those functions $f$ for which $\sup_{x \in \Xset} |f(x)| < \infty$.
+	\item For functions $f,g: \Xset \to \Reals$, we write $\dotp{f}{g}_n := \frac{1}{n}\sum_{i = 1}^{n} f(X_i) g(X_i)$, which we refer to as the \emph{empirical inner product}, and $\norm{f}_n^2 = \dotp{f}{f}_n$, which we refer to as the (squared) \emph{empirical norm}.
 For vectors $v \in \Reals^n$, we write $\norm{v}_p^p := \sum_{i = 1}^{n} \abs{v_i}^p$.
-	\item For a graph $G$, we use $\Lap_G$ to denote the Laplacian of $G$.
+When there is no chance of being confused, we will sometimes associate a function $f$ with its evaluations $(f(X_1),\ldots,f(X_n)) \in \Reals^n$.
+	\item In general, for a graph $G$ we use $\Lap_G$ to denote the Laplacian of $G$.
 The matrix $\Lap_G$ is a square, symmetric, positive semi-definite matrix, with $n$ eigenvalues $\lambda_k(G)$ and eigenvector $v_k(G)$ pairs, defined according to
@@
 	\end{equation*}
-	and ordered so that $0 \leq \lambda_1(G) \leq \ldots \leq \lambda_n(G)$. For convenience, when dealing with the neighborhood graph $G_{n,r}$ we denote $\Lap_{n,r} = \Lap_{G_{n,r}}$. 
+	and ordered so that $0 \leq \lambda_1(G) \leq \ldots \leq \lambda_n(G)$. As we've already seen, to reduce notational clutter we write $\Lap_{n,r}$ instead of $\Lap_{G_{n,r}}$. 
+	\item For sequences $a_n, b_n$, we write $a_n \lesssim b_n$ when there exists a constant $C > 0$ such that $a_n \leq C b_n$ for all $n$, and $a_n \asymp b_n$ when $a_n \lesssim b_n$ and $b_n \lesssim a_n$.
 	\item When convenient we write $\mathbf{X} = (X_1,\ldots,X_n)$. 
@@
 
-Before we get to our main results, we will review some of the relevant literature on minimax optimal nonparametric estimation and testing of Sobolev functions.
-We will also briefly recall some (of the many) estimators and tests which achieve these rates: in particular the \emph{Fourier series projection} and \emph{smoothing spline} methods.
-As we will see, Laplacian eigenmaps and Laplacian smoothing can be seen as graph-based, i.e.
+Before we get to our main results, we will review some of the relevant literature on minimax optimal estimation and testing of Sobolev functions.
+We will also recall some (of the many) estimators and tests which achieve these rates: in particular the \emph{Fourier series projection} and \emph{smoothing spline} methods.
+As we will see, one can interpret Laplacian eigenmaps and Laplacian smoothing as graph-based, i.e.
 discrete, counterparts to these ``continuum'' methods.
-
-We start by giving a (very) brief exposition of Sobolev spaces, with an emphasis on the material relevant in the statistical context.
 
@@
 
+We start by briefly introducing Sobolev spaces, emphasizing only the material relevant to our statistical context.
 Roughly speaking, Sobolev spaces contain functions $f \in \Leb^p(\Xset)$ with derivatives which themselves belong to $\Leb^p(\Xset)$.
-More formally, for given integers $s$ and $p > 0$, the Sobolev space $W^{s,p}(\Xset)$ consists of all functions $f \in \Leb^p(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^p(\Xset)$.
-The Sobolev-$\{s,p\}$ seminorm $\seminorm{f}_{W^{s,p}(\Xset)}$ and norm $\norm{f}_{W^{s,p(\Xset)}}$ are then given by
+More formally, for given integers $s$ and $p > 0$, the Sobolev space $W^{s,p}(\Xset)$ consists of all functions $f \in \Leb^p(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ exists and belongs to $\mathcal{L}^p(\Xset)$.
+For such functions, the $(s,p)$-Sobolev seminorm $\seminorm{f}_{W^{s,p}(\Xset)}$ and norm $\norm{f}_{W^{s,p(\Xset)}}$ are given by
 \begin{equation*}
@@
 \end{equation*}
-and for a given $M > 0$, the corresponding ball is $W^{s,p}(\Xset, M) = \set{f: \norm{f}_{W^{s,p}(\Xset)} \leq M}$.
+and for a given $M > 0$, the $(s,p)$-Sobolev ball is $W^{s,p}(\Xset, M) = \set{f: \norm{f}_{W^{s,p}(\Xset)} \leq M}$.
 In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space, and we adopt the usual convention of writing $H^s(\Xset) = W^{s,p}(\Xset)$ and $H^s(\Xset,M) = W^{s,2}(\Xset,M)$; we will confine our attention to this case hereafter.
-One special fact worth mentioning: when $\Xset$ is sufficiently smooth, for all Sobolev spaces $H^s(\Xset)$ the Holder space $C^{\infty}(\Xset)$ is dense in $H^s(\Xset)$.
-Said differently, for each integer $s$, the Sobolev space $H^s(\Xset)$ is simply the completion of $C^{\infty}(\Xset)$ in the norm $\norm{\cdot}_{H^s(\Xset)}$.
 
@@
 \end{equation*}
-To see the connection between $H^s(M)$ and $\Theta_s(M)$, for the moment let us fix the domain $\Xset = [0,1]^d$, and write $\set{\phi_k}_{k \in \mathbb{Z}^d}$ for the tensor product trigonometric polynomial basis of $\Leb^2(\Xset)$.
+To see the connection between $H^s(M)$ and $\Theta_s(M)$, for the moment let us fix the domain $\Xset = [0,1]^d$, and write $\set{\phi_k}_{k \in \mathbb{Z}^d}$ for the trigonometric basis of $\Leb^2(\Xset)$.
 Then the periodic Sobolev balls
@@
 \end{equation*}
-satisfy $H_{\textrm{per}}^s(\Xset,M) \subseteq H^s(\Xset,M')$ for a sufficiently large value of $M_1 > 0$.
+satisfy $H_{\textrm{per}}^s(\Xset,M) \subseteq H^s(\Xset,M_1)$ for a sufficiently large value of $M_1 > 0$.
 On the other hand, any $1$-periodic function $f \in H^s(\Xset,M)$ also belongs to $H_{\textrm{per}}^s(\Xset,M_0)$ for a sufficiently large value of $M_0 > 0$.
@@
 
-\textcolor{red}{(TODO 1)}: Refer also to the Sobolev embedding theorem, for at least three reasons: (1) to facilitate later discussion of why smoothing splines are ill-posed when $2s < d$; (2) to make explicit why the statistical problem isn't ill-posed when $2s < d$ in the random design setting; (3) to make clear why the testing problem is hopeless when $4s < d$.
-
-\subsection{Nonparametric regression: Fourier series and smoothing splines.}
+\textcolor{red}{(TODO 1)}: Introduce the Sobolev embedding theorem, for at least three reasons: (1) to facilitate later discussion of why smoothing splines are ill-posed when $2s < d$; (2) to make explicit why the statistical problem isn't ill-posed when $2s < d$ in the random design setting; (3) to make clear why the testing problem is hopeless when $4s < d$.
+Conclude with one special fact worth mentioning: regardless of $s$ and $d$, when $\Xset$ is sufficiently smooth for any Sobolev spaces $H^s(\Xset)$ the set of infinitely differentiable functions $C^{\infty}(\Xset)$ is dense in $H^s(\Xset)$.
+Said differently, for each integer $s$, the Sobolev space $H^s(\Xset)$ is simply the completion of $C^{\infty}(\Xset)$ in the norm $\norm{\cdot}_{H^s(\Xset)}$.
+
+\subsection{Nonparametric regression: Fourier series and smoothing splines}
 \label{subsec:continuum_methods}
-When the regression function $f_0$ is assumed to belong to a Sobolev space, two approaches particularly well-suited for nonparametric regression are orthogonal projection onto the Fourier basis, and smoothing splines.
-In the former, we estimate the Fourier coefficients $\wh{\theta}_k := \dotp{Y}{\phi_k}_n$, with the resultant function estimate $\wh{f}_{\OS}$ defined as
+Orthogonal projection onto the Fourier basis and smoothing splines are two approaches particularly well-suited for nonparametric regression over Sobolev spaces.
+In the former approach, we estimate the Fourier coefficients $\wh{\theta}_k := \dotp{Y}{\phi_k}_n$, with the resulting function estimate
 \begin{equation}
@@
 \end{equation}
-where $\mathrm{div}$ stands for the divergence, and $\nabla$ for the gradient.
+where $\mathrm{div}(\cdot)$ stands for the divergence, and $\nabla$ for the gradient.
 As mentioned previously, when the radius $r \to 0$ as $n \to \infty$ at an appropriate rate, the graph Laplacian $\Lap_{n,r}$ in fact converges to a weighted Laplace operator $\Delta_P$, given by
@@
 \label{eqn:laplace_operator_weighted}
-\Delta_{P} g := -\frac{1}{p}\mathrm{div}(p^2 \nabla g),
-\end{equation}
-with the eigenvectors $v_k(G_{n,r})$ converging in an appropriate sense to eigenfunctions of $\Delta_{P}$.
-The upshot is that when $P$ is the uniform measure---and thus the eigenfunctions of $\Delta_P = \Delta$ are in fact the trigonometric basis functions $\phi_k$---for an appropriate sequence of radii $r(n) \to 0$ and any fixed $\kappa \geq 0$ the estimator $\wh{f}_{\OS}$ is in fact the limit of $\wh{f}_{\LE}$, in the sense that
+\Delta_{P} g := -\frac{1}{p}\mathrm{div}(p^2 \nabla g);
+\end{equation}
+likewise the eigenvectors $v_k(G_{n,r})$ converge to eigenfunctions of $\Delta_{P}$.
+The upshot is that when $P$ is the uniform measure on $\Xset = [0,1]^d$---and thus the eigenfunctions of $\Delta_P = \Delta$ are precisely the trigonometric basis functions $\phi_k$--- and $\kappa \geq 0$ is fixed, the estimator $\wh{f}_{\OS}$ is in fact the limit of $\wh{f}_{\LE}$, in the sense that
 \begin{equation}
 \label{eqn:laplacian_eigenmaps_consistency}
-\abs{\wh{f}_{\LE}(X_i) - \wh{f}_{\OS}(X_i)} \overset{P}{\to} 0~~\textrm{as $n \to \infty$.}
+\abs{\wh{f}_{\LE}(X_i) - \wh{f}_{\OS}(X_i)} \overset{P}{\to} 0~~\textrm{as $n \to \infty$ and $r(n) \to 0$,}
 \end{equation} 
@@
 
-By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can also tie the smoothing spline estimator $\wh{f}_{\SM}$ to the Laplacian smoothing estimator $\wh{f}_{\LS}$.
-Using the definition~\eqref{eqn:laplace_operator}, recursively applying integration by parts, and then invoking the aforementioned convergence properties of $\Lap_{n,r}$, it can be shown that for any $f \in C^{2s}(\Xset)$,
+By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can also tie the penalty in the smoothing spline estimator $\wh{f}_{\SM}$ to that of the Laplacian smoothing estimator $\wh{f}_{\LS}$.
+Using the definition~\eqref{eqn:laplace_operator}, repeated integrating by parts, and then invoking the aforementioned convergence properties of $\Lap_{n,r}$, it can be shown that for any $f \in C^{2s}(\Xset)$ (see \citet{zhou11} for an argument in this spirit),
 \begin{equation}
 \label{eqn:pointwise_seminorm_convergence}
-\abs{f}_{H^s(\Xset)} = \int_{\Xset} \bigl(\Delta_P^sf\bigr)(x) f(x) \,dx = \lim_{n \to \infty} \frac{1}{n^{s + 1}r^{s(d + 2)}}f^T \Lap_{n,r}^s f.
-\end{equation}
-(See \citet{zhou11} for an argument along these lines.)
-While this does not imply a formal statement about the convergence of $\wh{f}_{\LS}$ to $\wh{f}_{SM}$ akin to~\eqref{eqn:laplacian_eigenmaps_consistency}---note, for instance, the difference between the domain of minimization in~\eqref{eqn:smoothing_spline} and the regularity required for~\eqref{eqn:pointwise_seminorm_convergence}---it nevertheless makes the connection between $\wh{f}_{\LS}$ and $\wh{f}_{\SM}$ obvious.
+\abs{f}_{H^s(\Xset)} = \int_{\Xset} \bigl(\Delta_P^sf\bigr)(x) \cdot f(x) \,dx = \lim_{n \to \infty} \frac{1}{n^{s + 1}r^{s(d + 2)}}f^T \Lap_{n,r}^s f.
+\end{equation}
+This does not imply a formal statement about the convergence of $\wh{f}_{\LS}$ to $\wh{f}_{\SM}$ along the lines of~\eqref{eqn:laplacian_eigenmaps_consistency}.
+Among other issues, the convergence in~\eqref{eqn:pointwise_seminorm_convergence} is over $f \in C^{2s}(\Xset)$ while the domain of minimization in~\eqref{eqn:smoothing_spline} is $H^{s}(\Xset)$.
+Nevertheless, it establishes a general connection between $\wh{f}_{\LS}$ and $\wh{f}_{\SM}$.
 
@@
 In particular, these estimators and tests are known to have error rates within a constant factor of minimax optimal.
+Throughout this section, we treat $M > 0$ as a fixed number (i.e.
+not varying with $n$.)
 
 \paragraph{Estimation rates.}
-Under appropriate regularity conditions, the minimax estimation rate over Sobolev balls is 
+The minimax estimation rate over Sobolev balls is 
 \begin{equation}
@@
 \end{equation}
-where the infimum is over all measurable maps $\wh{f}$, and the function class $\mc{H}^s(\Xset,M)$ is any one of $H^s(\Xset,M)$, $H_{\mathrm{per}}^s(\Xset,M)$ or $H_0^s(\Xset,M)$.
-Fourier series and smoothing spline estimators are two (of the many) estimators which certify the upper bound in~\eqref{eqn:sobolev_space_estimation_minimax_rate}.
-In the former case, if we assume that
+where the function class $\mc{H}^s(\Xset,M)$ is any one of $H^s(\Xset,M)$, $H_{\mathrm{per}}^s(\Xset,M)$ or $H_0^s(\Xset,M)$.
+
+Fourier series and smoothing spline estimators are two (of the many) estimators which certify the upper bound in~\eqref{eqn:sobolev_space_estimation_minimax_rate}. More specifically, if we assume that
 \begin{enumerate}[label=(P\arabic*)]
@@
 	\label{asmp:uniform_density}
-	the distribution $P$ is the uniform probability measure on the unit cube $\Xset = [0,1]^d$.
+	the distribution $P$ is the uniform probability measure on the unit cube $\Xset = [0,1]^d$,
 \end{enumerate}
 then $\wh{f}_{\OS}$ is minimax optimal over the Sobolev ellipsoids $H_{\mathrm{per}}^s(\Xset, M)$, meaning
-\begin{equation*}
-\sup_{f \in H_{\mathrm{per}}^s(\Xset,M)}\Ebb\Bigl[\norm{\wh{f}_{\OS} - f}_{L^2(\Xset)}^2\Bigr]  \lesssim n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d \geq 0$.}
-\end{equation*}
-An advantage of $\wt{f}_{\SM}$ over $\wh{f}_{\OS}$ is that it requires weaker assumptions on the design density in order to be minimax optimal.
-Instead of Assumption~\ref{asmp:uniform_density}, we will merely assume that the density is bounded above and below.
+\begin{equation}
+\label{eqn:fourier_series_estimation_rate}
+\sup_{f_0 \in H_{\mathrm{per}}^s(\Xset,M)}\Ebb\Bigl[\norm{\wh{f}_{\OS} - f_0}_{L^2(\Xset)}^2\Bigr]  \lesssim n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d \geq 0$.}
+\end{equation}
+In contrast, if we merely assume that
 \begin{enumerate}[label=(P\arabic*)]
@@
 	\end{equation*}
-	for all $x \in \Xset$.
+	for all $x \in \Xset$,
 \end{enumerate}
-The assumption~\ref{asmp:bounded_density} is itself a standard but significant assumption, but clearly it is much less stringent than~\ref{asmp:uniform_density}: it does not specify that either the density $p$ or domain $\Xset$ be known.
-Assuming~\ref{asmp:bounded_density}, the estimator $\wt{f}_{\SM}$ satisfies
-\begin{equation*}
+then $\wt{f}_{\SM}$ is minimax optimal over the Sobolev balls $H^s(\Xset,M)$, meaning
+\begin{equation}
+\label{eqn:smoothing_spline_estimation_rate}
 \sup_{f_0 \in H^s(\Xset,M)}\Ebb\Bigl[\norm{\wh{f}_{\SM} - f_0}_{L^2(\Xset)}^2\Bigr]  \lesssim n^{-2s/(2s + d)}~~\textrm{for all $2s > d$.}
-\end{equation*}
-The restriction $2s > d$ is a severe limitation of smoothing splines relative to  orthogonal series estimators.
-It is worth emphasizing that this is a fundamental defect of smoothing splines, as opposed to any flaw in the theory; when $2s \leq d$, the optimization problem~\eqref{eqn:smoothing_spline} is not even well-defined!
-
-\textcolor{red}{(TODO 2):} succinctly explain why, by referring to the Sobolev Embedding Theorem.
+\end{equation}
+These results reveal some relative strengths and weaknesses of $\wh{f}_{\OS}$ and $\wh{f}_{\SM}$.
+On the one hand, the result for smoothing splines clearly invokes weaker regularity conditions than the result for Fourier series projection.
+The assumption~\ref{asmp:bounded_density} is less stringent than~\ref{asmp:uniform_density}, in that it does not specify that either the density $p$ or domain $\Xset$ be known.
+Moreover,~\eqref{eqn:fourier_series_estimation_rate} requires that the regression function $f_0$ be periodic, whereas~\eqref{eqn:smoothing_spline_estimation_rate} does not.
+On the other hand, the restriction $2s > d$ in~\eqref{eqn:smoothing_spline_estimation_rate} is a severe limitation of smoothing splines relative to Fourier series estimators.
+It is worth emphasizing that this is a fundamental defect of smoothing splines, as opposed to any flaw in the theory; when $2s \leq d$ the optimization problem~\eqref{eqn:smoothing_spline} is not even well-defined!
+Interestingly, many of these points will reappear when we turn to graph-based estimators.
+To be explicit: we will see that the Laplacian smoother $\wh{f}_{\LS}$ fails when $d$ is too large relative to $s$; that our analysis of $\wh{f}_{\LE}$ requires that $f_0$ be periodic when $s > 1$; and perhaps most notably, that neither the estimators nor our analysis requires $p$ or $\Xset$ to be known.
+
+\textcolor{red}{(TODO 2):} succinctly explain why smoothing splines are ill-posed when $2s \leq d$, by referring to the Sobolev Embedding Theorem.
 
@@
 
-In the nonparametric regression goodness-of-fit testing problem, we ask for a test function (formally, a Borel measurable map $\phi: \mc{X}^n \times \Reals^n \to \{0,1\}$) which can distinguish between the null and alternative hypotheses
+In the goodness-of-fit testing problem, we ask for a test function (formally, a Borel measurable map $\phi: \mc{X}^n \times \Reals^n \to \{0,1\}$) which can distinguish between the null and alternative hypotheses
 \begin{equation}
@@
 \end{equation}
-We measure the performance of a test $\phi$ by its Type I error $\Ebb_0[1 - \phi]$, and Type II error $\mathbb{E}_{f_0}[\phi]$.
+We measure the performance of a test $\phi$ by its Type I error $\Ebb_0[\phi]$, and Type II error $\mathbb{E}_{f_0}[1 - \phi]$.
 It is well known that no test $\phi$ with controlled Type I error---say, $\Ebb_0[\phi] \leq 1/b$ for some constant $b > 1$---
@@
 As a result, it is standard to impose two kinds of restrictions on the alternative hypothesis class $\mathbf{H}_a$.
-The first is a smoothness condition on $f_0$: in our case, we will assume that $f_0$ belongs to a Sobolev ball $\mc{H}^s(\Xset,M)$.
-The latter is a separation condition, enforcing a minimum distance between $f_0$ and $0$: in our case, we will insist that $\norm{f_0  - 0}_{\Leb^2(\Xset)} > \epsilon$ for some $\epsilon > 0$.
+The first is a smoothness condition on $f_0$: in our case, we will assume that $f_0$ belongs to a Sobolev ball $H^s(\Xset,M)$.
+The latter is a separation condition enforcing a minimum distance between $f_0$ and $0$: in our case, we will insist that $\norm{f_0  - 0}_{\Leb^2(\Xset)} > \epsilon$ for some $\epsilon > 0$.
 
@@
 \label{eqn:sobolev_space_testing_critical_radius}
-\epsilon_n^{\star}\Bigl(\mc{H}^s(\Xset,M)\Bigr):= \inf\biggl\{\epsilon: \inf_{\phi} \Bigl\{\Ebb_0[\phi] +  \sup_{\substack{f_0 \in \mc{H}^s(\Xset,M), \\ \norm{f_0}_{\Leb^2(\Xset)} \geq \epsilon}} \Ebb_{f_0}[1 - \phi]\Bigr\} \leq \frac{2}{b} \biggr\} \asymp n^{-2s/(4s + d)}
-\end{equation}
-for any $4s > d$, and $\mc{H}^s(\Xset,M)$ any one of $H^s(\Xset,M)$, $H_{\mathrm{per}}^s(\Xset,M)$ or $H_0^s(\Xset,M)$.
+\epsilon_n^{\star}\Bigl(\mc{H}^s(\Xset,M)\Bigr):= \inf\biggl\{\epsilon: \inf_{\phi} \Bigl\{\Ebb_0[\phi] +  \sup_{\substack{f_0 \in \mc{H}^s(\Xset,M), \\ \norm{f_0}_{\Leb^2(\Xset)} \geq \epsilon}} \Ebb_{f_0}[1 - \phi]\Bigr\} \leq \frac{2}{b} \biggr\} \asymp n^{-2s/(4s + d)}~~\textrm{for all $4s > d$,}
+\end{equation}
+where $\mc{H}^s(\Xset,M)$ is any one of $H^s(\Xset,M)$, $H_{\mathrm{per}}^s(\Xset,M)$ or $H_0^s(\Xset,M)$.
 Just as $\wh{f}_{\OS}$ and $\wh{f}_{\SM}$ achieve minimax optimal estimation rates, tests using the plug-in statistics
@@
 \end{equation*}
-are minimax optimal for the goodness-of-fit testing problem over Sobolev balls; see \citet{ingster2009} for analysis of the former statistic, and \citet{liu2019} for analysis of the latter (when $2s > d$).
+are minimax rate-optimal for the goodness-of-fit testing problem over Sobolev balls; see \citet{ingster2009} for analysis of the former statistic, and \citet{liu2019} for analysis of the latter (when $2s > d$).
 
@@
 Suppose that $\Xset$ satisfies the manifold hypothesis, meaning informally that it has intrinsic dimension $m < d$ (see~\ref{asmp:domain_manifold} in Section~\ref{sec:graph_sobolev_classes} for a formal definition).
-In this context, both~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} now hold with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
+In this context, both~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
 What's more, suitable adaptations of the estimators $\wh{f}_{\OS}$ and $\wt{f}_{\SM}$, and the test statistics $T_{\OS}$ and $T_{\SM}$, achieve these faster rates.
-We will not further review the details here, except to emphasize one point: in this context, in order for either the smoothing spline or Fourier series approaches to be optimal, the domain $\Xset$ must be known; as we will see, our graph-based procedures will be optimal without this requirement.
-We will return to this topic, and comment more generally on how different methods and analyzes make use of different assumptions on the design distribution $P$, in Section~\ref{sec:discussion}.
+We will not further review the details here, except to emphasize one point: unlike in the full-dimensional context, in order for either the smoothing spline or Fourier series approaches to optimally leverage the manifold hypothesis, the domain $\Xset$ must be known.
+In contrast, we will see that our graph-based procedures will be optimal without this knowledge; in this sense, they automatically ``feel'' the geometry of $\Xset$.
+We will return to this point in Section~\ref{sec:discussion}.
 
@@
 
+With the relevant function spaces and minimax rates established, we can now summarize our results.
 At a high level, the connections between the graph-based methods~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} and the continuum methods outlined in Subsection~\ref{subsec:continuum_methods} give us reason to hope that the strong statistical properties of the latter are shared by the former.
-Making this rough statement precise is the content of this paper.
-In particular, we will prove the following results:
+Making this rough statement precise is the major contribution of this paper.
+In particular, we will prove the following.
 
 \begin{itemize}
-	\item \textbf{Minimax optimal estimation.}
-	With high probability, $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2/(2 + d)}$ over all $f_0 \in H^1(\Xset;M)$. The same conclusions hold with respect to $\wt{f}_{\LS}$ when $d \leq 4$. 
+	\item \textbf{Minimax optimal estimation.} With high probability, $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2/(2 + d)}$ over all $f_0 \in H^1(\Xset;M)$.
+The same conclusions hold with respect to $\wt{f}_{\LS}$ when $d \leq 4$.
 	\item \textbf{Minimax optimal testing.}
@@
 	\item \textbf{Taking advantage of higher-order smoothness conditions.} Under appropriate boundary conditions on the function classes $H^s(\Xset)$ and additional smoothness conditions on the density $p$, with high probability the error $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2s/(2s + d)}$ over all $f_0 \in H^s(\Xset;M)$.
-Additionally, a test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in H^s(\Xset;M)$ satisfies $\norm{f_0}_{\Leb^2(\Xset)} \gtrsim \max\{n^{-2s/(4s + d)}\}$ and $d < 4$.
+Additionally, a test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in H^s(\Xset;M)$ satisfies $\norm{f_0}_{\Leb^2(\Xset)} \gtrsim n^{-2s/(4s + d)}$ and $d < 4$.
 	\item \textbf{Manifold adaptivity.}
@@
 \end{itemize}
+
+Each of these results rest on various properties of graph Sobolev classes, which we now turn to.
 
@@
 To understand the relevance of the graph Sobolev balls to the analysis of our graph-based procedures, let us return for a moment to the continuum setting.
-The statistical analysis of continuous-time methods such as Fourier series projection and smoothing splines rests on two facts about the continuum classes $\mc{H}^s(\Xset)$: first, an \textit{a priori} assumption that the regression function $f_0$ belongs to an appropriate Sobolev ball $\mc{H}(\Xset,M)$, and second that the asymptotic decay of the eigenvalues of the continuum Laplace operator $\Delta$ obey Weyl's law, meaning $\lambda_k(\Delta) \asymp \abs{k}^{2s}$ for $k \in \mathbb{Z}^d$.
+Upper bounds on the risk of continuous-time methods such as Fourier series projection and smoothing splines rests on two facts about the continuum classes $\mc{H}^s(\Xset)$: first, an \textit{a priori} assumption that the regression function $f_0$ belongs to an appropriate Sobolev ball $\mc{H}^{s}(\Xset,M)$, and second that the asymptotic decay of the eigenvalues of the continuum Laplace operator $\Delta$ obey Weyl's law, meaning $\lambda_k(\Delta) \asymp \abs{k}^{2s}$ for $k \in \mathbb{Z}^d$.
 In this section, we will show that both these statements have rough analogues in the graph setting.
@@
 \label{eqn:graph_sobolev_seminorm_asymptotic}
-f^T \Lap_{n,r}^s f \lesssim M(n^{s + 1}r^{s(d + 2)})
+f^T \Lap_{n,r}^s f \lesssim M(n^{s + 1}r^{s(d + 2)}),
 \end{equation}
@@
 \label{subsec:graph_sobolev_seminorm}
-We will state and prove three different results involving the graph Sobolev semi-norm $f_0^T \Lap_{n,r}^s f_0$.
+We will state and prove three different results involving the graph Sobolev semi-norm $f^T \Lap_{n,r}^s f$.
 In each case, we will derive that a version of the asymptotic relation~\eqref{eqn:graph_sobolev_seminorm_asymptotic} holds with high probability, whenever $r \to 0$ sufficiently slowly as $n \to \infty$.
@@
 	\label{asmp:domain}
-	\textcolor{red}{(TODO): Explain what a Lipschitz domain is.} The density $p$ is supported on a Lipschitz domain $\Xset$.
+	The density $p$ is supported on a Lipschitz domain $\Xset$. \textcolor{red}{(TODO 3)}: Formally define a Lipschitz domain.
 \end{enumerate}
@@
 The proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm}, as with all of our proofs, is contained in the Appendix.
-There are a few remarks worth making regarding Lemma~\ref{lem:first_order_graph_sobolev_seminorm}, but they all apply equally well to our analysis of $f^T\Lap_{{n,r}}^sf$ for $s > 1$, and so will wait to make them.
+There are a few remarks worth making regarding Lemma~\ref{lem:first_order_graph_sobolev_seminorm}, but they all apply equally well to our analysis of $f^T\Lap_{{n,r}}^sf$ for $s > 1$, and so we will wait to make them.
 
@@
 Let $r = r(n)$ be a sequence of connectivity radii satisfying $r(n) \to 0$ and $r(n) \gtrsim n^{-1/(2(s - 1) +d)}$ as $n \to \infty$.
-Then for all $n$ sufficiently large, there exists a constant $C_2 > 0$ which depends only on $\Xset$, $d$, $p_{\max}$ and $K$ such that
+Then there exists a constant $C_2 > 0$ which depends only on $\Xset$, $d$, $p_{\max}$ and $K$ such that for all $n$ sufficiently large,
 	\begin{equation}
@@
 	\item 
-	The scaling (in $n$ and $r$) on the right hand side of~\eqref{eqn:graph_sobolev_seminorm} will be familiar to the informed reader.
-For a sufficiently regular function $f \in C^3(\Xset,M)$ and density $p \in C^1(\Xset,p_{\max})$, it is known \textcolor{red}{(references)} that $\frac{1}{nr^{d + 2}}\Lap_{n,r}f \to \Delta_pf$ for $n^{-1/(2 + d)}$.
-A recursive argument shows that when $f \in C^{2s + 1}(\Xset)$, $p \in C^{2s}(\Xset)$, and $r$ is sufficiently large,
-	\begin{equation}
-	\label{eqn:graph_sobolev_seminorm_heuristic}
-	\Bigl(\frac{1}{nr^{d + 2}}\Lap_{n,r}\Bigr)^s f \to \Delta_p^sf \Longrightarrow \frac{1}{n^{s + 1} r^{s(d + 2)}} f^T \Lap_{n,r}^s f \to \int_{\Xset} f(x) \cdot  \Delta_p^sf(x) p(x) \,dx \leq c M^2.
-	\end{equation}
-
-	\textcolor{red}{(TODO 1)}: Figure out the right scaling of $r$ for this heuristic argument to go through.
-
-	In fact, the actual proof of Theorem~\ref{thm:graph_sobolev_seminorm} is rather more involved.
-Among other subtleties we highlight the mismatch, in the heuristic argument just given, between the number of smooth derivatives $f$ possesses ($2s + 1$) and the degree of the Laplacian operator $\Lap_{{n,r}}^s$ $(s)$; this is unsuitable for our purposes.
-We use a more careful analysis to prove Theorem~\eqref{thm:graph_sobolev_seminorm}, but note well that~\eqref{eqn:graph_sobolev_seminorm} is only an upper bound rather than a statement about consistency.
+	The inequality~\eqref{eqn:graph_sobolev_seminorm} resembles~\eqref{eqn:pointwise_seminorm_convergence}.
+However, as we pointed out previously to derive~\eqref{eqn:pointwise_seminorm_convergence} one needs $f \in C^{2s}(\Xset,M)$ as opposed to $f \in H^s(\Xset,M)$, and imposing this extra degree of regularity is unsuitable for our purposes.
+We use a considerably more careful analysis to prove Theorem~\eqref{thm:graph_sobolev_seminorm}, but note well that~\eqref{eqn:graph_sobolev_seminorm} is only an upper bound rather than a statement about consistency.
 This is the price we pay for insisting that the assumed number of smooth derivatives match the degree of the Laplacian; fortunately an upper bound will be sufficient for our purposes in Section~\ref{sec:minimax_optimal_graph_Laplacian_methods}.
@@
 	\item In~\eqref{eqn:first_order_graph_sobolev_seminorm} and~\eqref{eqn:graph_sobolev_seminorm}, we state the upper bound as a function of the Sobolev norm $\norm{f}_{H^s(\Xset)}$ of $f$, rather than the Sobolev seminorm $|f|_{H^s(\Xset)}$.
-The derivations in~\eqref{eqn:graph_sobolev_seminorm_heuristic} might leave us some hope that it suffices to assume a bounded seminorm.
-Indeed, in Lemma~\ref{lem:first_order_graph_sobolev_seminorm}, we may relax this---that is, we may replace $M$ by $|f|_{H^1(\Xset)}$---because the graph Laplacian $\Lap_{n,r}$ annihilates the constant part of any function $f$.
-In the latter case, we may not replace $M$ by $|f|_{H^s(\Xset)}$.
-
-	\textcolor{red}{(TODO 2)}: Explain why.
-
-	\textcolor{red}{(TODO 3)}: Check with Ryan and Siva--- or if not them, Google--- to figure out whether boundedness in seminorm plus zero-trace condition implies boundedness in norm.
-Intuitively, it feels like it does.
-
-	\textcolor{red}{(TODO 4)}: Go back and check whether Fourier series projection requires boundedness in norm, or only in semi-norm.
-	\item We note also that~\eqref{eqn:first_order_graph_sobolev_seminorm} and~\eqref{eqn:graph_sobolev_seminorm}---which each hold with probability $1 - \delta$---are proportional to $1/\delta$, as opposed to $\log(1/\delta)$ as is typical of exponential concentration.
+The derivations in~\eqref{eqn:pointwise_seminorm_convergence} might leave us some hope that assuming a bounded seminorm is enough.
+Indeed, in Lemma~\ref{lem:first_order_graph_sobolev_seminorm} we can replace $M$ by $|f|_{H^1(\Xset)}$, because the graph Laplacian $\Lap_{n,r}$ annihilates the constant part of any function $f$.
+However in the latter case, we may not replace $M$ by $|f|_{H^s(\Xset)}$.
+
+	\textcolor{red}{(TODO 4)}: Explain why.
+
+	\textcolor{red}{(TODO 5)}: Check with Ryan and Siva--- or if not them, Google--- to figure out whether boundedness in seminorm plus zero-trace condition implies boundedness in norm.
+Intuitively, it feels like it does: maybe this is just Poincare's inequality?
+
+	\textcolor{red}{(TODO 6)}: Go back and check whether Fourier series projection requires boundedness in norm, or only in semi-norm.
+	\item We note also that~\eqref{eqn:first_order_graph_sobolev_seminorm} and~\eqref{eqn:graph_sobolev_seminorm}---which each hold with probability $1 - \delta$---are proportional to $1/\delta$, as opposed to the tighter bound of $\log(1/\delta)$.
 The reason: the Sobolev assumption $f \in H^s(\Xset,M)$ only grants us control over the second moment of $f$ and its derivatives, and so we are afforded only the relatively weak concentration implied by Markov's inequality.
-Under higher-order moment control on $f$ and its derivatives---for instance if $f \in C^s(\Xset,M)$---we can get sharper upper bounds.
+If we assume $f$ and its derivatives are more uniformly bounded--for instance if $f \in C^s(\Xset,M)$---we can get sharper upper bounds.
 \end{itemize}
 
-Finally, we note that while the particular requirement in Theorem~\ref{thm:graph_sobolev_seminorm} that $f$ be zero-trace may be a tad overkill, it is essential that $f$ display some type of periodicity for the upper bound~\eqref{eqn:graph_sobolev_seminorm} to hold.
-
-\textcolor{red}{(TODO 5)}: Elaborate on this: (a) reference developments in the literature showing that the eigenvectors of $\Lap_{n,r}$ satisfy Neumann boundary conditions in the limit $n \to \infty, r \to 0$, and (b) state, as a Lemma, that there exists a function $g$ which is not zero-trace, and for which the upper bound~\eqref{eqn:graph_sobolev_seminorm} is not satisfied.
+Finally, we should point out that while the particular requirement in Theorem~\ref{thm:graph_sobolev_seminorm} that $f$ be zero-trace may not be strictly necessary, it is essential that $f$ display some type of periodicity for the upper bound~\eqref{eqn:graph_sobolev_seminorm} to hold.
+
+\textcolor{red}{(TODO 7)}: Elaborate on this: (a) reference developments in the literature showing that the eigenvectors of $\Lap_{n,r}$ satisfy Neumann boundary conditions in the limit $n \to \infty, r(n) \to 0$, and (b) state, as a Lemma, that there exists a function $g$ which is not zero-trace, and for which the upper bound~\eqref{eqn:graph_sobolev_seminorm} is not satisfied.
 
@@
 
-As mentioned in the introduction, an attractive quality of the neighborhood graph is that it automatically captures the geometry of $P$ and $\Xset$.
+As mentioned in the introduction, an attractive quality of the neighborhood graph $G_{n,r}$ is that it captures the geometry of $P$ and $\Xset$.
 We now show one manifestation of this phenomenon: the graph Sobolev semi-norm automatically adapts to the intrinsic dimension of $\Xset$.
@@
 Let $r = r(n)$ be a sequence of connectivity radii satisfying $r(n) \to 0$ and $r(n) \gtrsim n^{-1/(2(s - 1) + m)}$ as $n \to \infty$.
-Then for all $n$ sufficiently large, there exists a constant $C_3 > 0$ which depends only on $\Xset, m$ and $p$ such that
+Then there exists a constant $C_3 > 0$ which depends only on $\Xset$, $m$, $p_{\max}$ and $K$ such that for all $n$ sufficiently large,
 	\begin{equation*}
@@
 \begin{itemize}
-	\item \textcolor{red}{(TODO 6)}: Explain why we require $s \leq 3$.
-	\item \textcolor{red}{(TODO 7)}: See if we can replace the Holder assumption $f \in C^s$ with the Sobolev assumption $f \in H^s$.
+	\item \textcolor{red}{(TODO 8)}: Explain why we require $s \leq 3$.
+	\item \textcolor{red}{(TODO 9)}: See if we can replace the Holder assumption $f \in C^s(\Xset,M)$ with the Sobolev assumption $f \in H^s(\Xset,M)$.
 \end{itemize}
@@
 Let $r = r(n)$ be a sequence of connectivity radii satisfying $r(n) \to 0$ as $n \to \infty$.
-Then for all $n$ sufficiently large, there exist constants $c_4$ and $C_4$ which depend only on $p_{\min}, p_{\max},d$ and $\Xset$, such that the following statement
+Then there exist constants $c_4$ and $C_4$ which depend only on $p_{\min}, p_{\max},d$ and $\Xset$ and $K$, such for all $n$ sufficiently large the following statement
 	\begin{equation}
@@
 Let $r = r(n)$ be a sequence of connectivity radii satisfying $r(n) \to 0$ as $n \to \infty$.
-Then there exist constants $c_5$ and $C_5$ which depend only on $p_{\min}, p_{\max}$,$m$ and $\Xset$, such that the following statement
+Then there exist constants $c_5$ and $C_5$ which depend only on $p_{\min}, p_{\max}$, $m$, $\Xset$, and $K$, such that for all $n$ sufficiently large the following statement
 	\begin{equation}
@@
 \end{lemma}
-
-The second term in the lower bound of~\eqref{eqn:neighborhood_eigenvalue} and~\eqref{eqn:neighborhood_eigenvalue_manifold} is an artifact of the well known inequality $\lambda_n(G) \leq 2 \max_{i = 1,\ldots,n} \mathbf{D}_{ii}$---in words, that the maximum eigenvalue of the Laplacian can never be more than twice the maximum degree of the graph.
-To obtain optimal rates for our Laplacian eigenmaps estimator and test statistic, we will need the eigenvalues $\lambda_k(G_{n,r})$ to asymptotically scale like~\eqref{eqn:neighborhood_graph_eigenvalue_asymptotic} for particular values of $k$.
+The upper bounds in~\eqref{eqn:neighborhood_eigenvalue} and~\eqref{eqn:neighborhood_eigenvalue_manifold} resemble~\eqref{eqn:neighborhood_graph_eigenvalue_asymptotic} as promised.
+However,  depending on the values of $k$ and $r$, the lower bounds may scale like $nr^d$ (or $nr^m$) as opposed to~\eqref{eqn:neighborhood_graph_eigenvalue_asymptotic}.
+This is an artifact of the well known inequality $\lambda_n(G_{n,r}) \leq 2 \max_{i = 1,\ldots,n} \mathbf{D}_{ii} $---in words, that the maximum eigenvalue of the Laplacian can never be more than twice the maximum degree of the graph.
+To obtain optimal rates for our Laplacian eigenmaps estimator and test statistic, we will need the eigenvalues $\lambda_k(G_{n,r})$ to asymptotically scale like~\eqref{eqn:neighborhood_graph_eigenvalue_asymptotic} for a particular sequence of $k = k(n)$.
 Doing some algebra, we deduce in Corollary~\ref{cor:neighborhood_eigenvalue} various conditions on $k$ and $r$ under which the graph eigenvalue scales at least as fast as is desired, i.e.
@@
 	\label{cor:neighborhood_eigenvalue}
-	Suppose the distribution $P$ satisfies~\ref{asmp:bounded_density} and~\ref{asmp:domain}, and the kernel $K$ satisfies~\ref{asmp:kernel}. Then each of the following statements:
+	Suppose the distribution $P$ satisfies~\ref{asmp:bounded_density} and~\ref{asmp:domain}, and the kernel $K$ satisfies~\ref{asmp:kernel}.
+Then for all $n$ sufficiently large each of the following statements:
 	\begin{itemize}
@@
 	\item We will use inequality~\eqref{eqn:neighborhood_eigenvalue_1} to upper bound the bias of the Laplacian eigenmaps estimator $\wh{f}_{\LE}$, and inequality~\eqref{eqn:neighborhood_eigenvalue_2} to upper bound the bias of the Laplacian eigenmaps test statistic $T_{\LE}$.
-	\item Let us briefly summarize the various restrictions on the asymptotic scaling of $r$.
+	\item It is worth summarizing the various restrictions we've made on the asymptotic scaling of $r$ in different parts of Section~\ref{sec:graph_sobolev_classes}.
 Corollary~\ref{cor:neighborhood_eigenvalue} tells us that the desired eigenvalue scalings~\eqref{eqn:neighborhood_graph_eigenvalue_asymptotic} kick in only when the neighborhood graph radius $r$ is sufficiently small.
-On the other hand, in order for the probabilistic bounds of this corollary to be non-facile, the neighborhood graph radius must be sufficiently large, roughly speaking $r \gtrsim (\log n/n)^{1/d}$ (or in the manifold case $r \gtrsim (\log n/n)^{1/m}$).
-What's more, the conclusions of Theorem~\ref{thm:graph_sobolev_seminorm} hold only when neighborhood graph radius $r$ is sufficiently large.
+On the other hand, in order for the probabilistic bounds of this corollary to be non-facile, the neighborhood graph radius must also be sufficiently large, roughly speaking $r \gtrsim (\log n/n)^{1/d}$ (or in the manifold case $r \gtrsim (\log n/n)^{1/m}$).
+What's more, the conclusions of Theorems~\ref{thm:graph_sobolev_seminorm} and~\ref{thm:manifold_graph_sobolev_seminorm} hold only when neighborhood graph radius $r$ is sufficiently large: $r \gtrsim n^{-1/(2(s - 1) + d)}$ in the former case, and $r \gtrsim n^{-1/(2(s - 1) + m)}$ in the latter case.
 Consequently, our results in Section~\ref{sec:minimax_optimal_graph_Laplacian_methods} regarding statistical optimality of Laplacian eigenmaps will hold only for the values of $r$ which satisfy all these restrictions, when such values exist.
@@
 
-We now turn to formalizing the main conclusions of this work: that in various settings and under various conditions, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and the resulting test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test}, are asymptotically minimax-rate optimal.
+We now turn to formalizing the main conclusions of this work: that in various settings and under various conditions, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and the resulting test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test}, are asymptotically minimax rate-optimal.
 We divide these conclusions into three parts.
@@
 \label{subsec:minimax_first_order_smoothness}
-Let us start by assuming $f_0 \in H^1(\Xset,M)$, and proving upper bounds on the risk of our various graph-based procedures.
+Let us start by assuming $f_0 \in H^1(\Xset,M)$, and proving upper bounds on the error of our various graph-based procedures.
 Pleasingly, these results will go through with essentially no additional regularity conditions---additional, that is, to the standard~\ref{asmp:bounded_density} and~\ref{asmp:domain}---on either the regression function $f_0$ or the density $p$.
@@
 In Theorem~\ref{thm:laplacian_smoothing_estimation1} we consider the Laplacian smoothing estimator $\wt{f}_{\LS}$, and draw a similar conclusion to Theorem~\ref{thm:laplacian_eigenmaps_estimation1}.
-One difference: to establish Theorem~\ref{thm:laplacian_smoothing_estimation1}, we will require (much) tighter restrictions on $r$ than were required for Theorem~\ref{thm:laplacian_eigenmaps_estimation1}.
+One difference: to establish Theorem~\ref{thm:laplacian_smoothing_estimation1}, we will impose a (much) tighter upper bound on $r$ than was required for Theorem~\ref{thm:laplacian_eigenmaps_estimation1}.
 \begin{enumerate}[label=(R\arabic*)]
@@
 	\begin{equation*}
-	\biggl(\frac{\log n}{n}\biggr)^{1/d} \lesssim r(n) \lesssim n^{-3/(4 + 2d)} M^{(d - 4)/(4 + 2d)}
-	\end{equation*}
-	for all $n$ sufficiently large.
+	\biggl(\frac{\log n}{n}\biggr)^{1/d} \lesssim r(n) \lesssim n^{-3/(4 + 2d)} M^{(d - 4)/(4 + 2d)}.
+	\end{equation*}
 \end{enumerate}
@@
 	\item The range of radii $r$ imposed by~\ref{asmp:le_kernel_radius_estimation} and~\ref{asmp:ls_kernel_radius_estimation} are compatible with practice, where by far the most common choice of radius is the connectivity threshold $r(n) \asymp (\log(n)/n)^{1/d}$, chosen to make the graph $G_{n,r}$ as sparse as possible while still being connected.
-	\item For Laplacian eigenmaps, the choice of $\kappa$ is made to balance the squared-bias and variance terms, i.e. chosen so that with high probability
+	\item For Laplacian eigenmaps, $\kappa$ is chosen to balance the squared-bias and variance terms, i.e. chosen as small as possible while still with high probability satisfying 
 	\begin{equation*}
@@
 	Unlike in the analysis of the Fourier series projection estimator, here the (conditional on $\mathbf{X}$) bias is itself a random variable.
-To upper bound it, we rely on the pointwise and spectral convergence of neighborhood graph Laplacian semi-norms and eigenvalues, as established in Section~\ref{sec:graph_sobolev_classes}.
-	\item A similar set of conclusions hold with respect to the choice of $\rho$ for the Laplacian smoothing estimator $\wt{f}_{\LS}$, where we seek to balance
+To upper bound it, we rely on the results of Section~\ref{sec:graph_sobolev_classes}.
+	\item Similarly, for Laplacian smoothing $\rho$ is chosen to balance squared-bias and variance, i.e. chosen as large as possible while still with high probability satisfying
 	\begin{equation*}
@@
 The second point is the source of the tighter upper bound on $r$ we require in~\ref{asmp:ls_kernel_radius_estimation}.
-	\item Our graph-based estimators are defined only at the design points $(X_1,\ldots,X_n)$ and we therefore measure error using the in-sample mean-squared error $\norm{\cdot}_n^2$.
+	\item Our graph-based estimators are defined only at the design points $X_1,\ldots,X_n$ and we therefore measure error using the empirical norm $\norm{\cdot}_n^2$.
 Because we work in the random design framework, the in-sample error is a random variable, and our bound is thus a bound in high probability.
@@
 Assuming that the extension of our estimators and the regression function $f_0$ are suitably smooth, tools from empirical process theory will guarantee that the $\Leb^2(\Xset)$ error is not too much greater than the in-sample error, but we do not further pursue the details here.
-	\item A key difference between Theorems~\ref{thm:laplacian_eigenmaps_estimation1} and~\ref{thm:laplacian_smoothing_estimation1} is that the latter holds only when $d \leq 4$.
+	\item A key difference between Theorems~\ref{thm:laplacian_eigenmaps_estimation1} and~\ref{thm:laplacian_smoothing_estimation1} is that the latter holds only when $d < 4$.
 This restriction reflects a fundamental drawback in (penalized) least squares estimators more generally.
@@
 We should also mention that in a related context (estimation over a $d$-dimensional lattice graph), \cite{sadhanala2016} arrive at the same restriction $d < 4$.
-As in that work, when $d = 4$ our analysis results in a bound on risk within a $\log n$ factor of the minimax risk (in our context, we set $r \asymp (\log n/n)^{1/4}$ to derive this rates.)
+As in that work, when $d = 4$ our analysis results in a bound on the empirical norm within a $\log n$ factor of the minimax risk (in our context, we would set $r \asymp (\log n/n)^{1/4}$, slightly outside the scaling~\ref{asmp:ls_kernel_radius_estimation}, to obtain this rate.)
 \end{itemize}
@@
 Then let the Laplacian eigenmaps test $\wh{\phi}_{\LE} := \1\bigl\{T_{\LE} \leq \wh{t}_b\bigr\}$, and the Laplacian smoothing test $\wt{\phi}_{\LS} := \1\bigl\{T_{\LS} \leq \wt{t}_b\bigr\}$; clearly the tests $\wh{\phi}_{\LE}$ and $\wt{\phi}_{\LS}$ depend on $b$, but we suppress this notationally.
-The number $b$---or more accurately $1/b$---reflects the tolerated Type I and Type II error.
+The number $b$ is selected by the user, with the choice made based on the tolerated level of Type I and Type II error.
 In fact, some standard computations show that the thresholds $\wh{t}_b$ ($\wt{t}_b$) are sufficiently large to control the Type I error of $\wh{\phi}_{\LE}$ ($\wt{\phi}_{\LS}$),
 \begin{equation*}
-\Ebb_0\bigl[\wh{\phi}_{\LE}\bigr],~~\Ebb_0\bigl[\wt{\phi}_{\LS}\bigr] \leq \frac{1}{b^2};
+\Ebb_0\Bigl[\wh{\phi}_{\LE}\Bigr],~~\Ebb_0\Bigl[\wt{\phi}_{\LS}\Bigr] \leq \frac{1}{b^2};
 \end{equation*}
 we relegate these computations to the appendix (see Lemmas~\ref{lem:le_fixed_graph_testing} and~\ref{lem:ls_fixed_graph_testing}).
-It remains to show that these tests have small Type II error whenever $f_0$ is separated from $0$ by more than the critical radius $\epsilon_n^{\star}(\mc{H}^s(\Xset,M))$, which we recall was defined in~\eqref{eqn:critical_radius}.
-We proceed in a parallel manner to Section~\ref{subsubsec:first_order_smoothness_estimation}, starting first with the Laplacian eigenmaps test and moving on to the Laplacian smoothing test.
+It remains to show that these tests have small Type II error whenever $f_0$ is separated from $0$ by more than the critical radius $\epsilon_n^{\star}(\mc{H}^s(\Xset,M))$, which we recall from~\eqref{eqn:sobolev_space_testing_critical_radius}.
+To show this, we proceed in a parallel manner to Section~\ref{subsubsec:first_order_smoothness_estimation}, starting first with the Laplacian eigenmaps test and moving on to the Laplacian smoothing test.
 
@@
 	\begin{equation*}
-	\biggl(\frac{\log n}{n}\biggr)^{1/d} \lesssim r(n) \lesssim M^{-4/(4 + d)}n^{-2/(4 + d)}
-	\end{equation*}
-	for all $n$ sufficiently large.
+	\biggl(\frac{\log n}{n}\biggr)^{1/d} \lesssim r(n) \lesssim M^{-4/(4 + d)}n^{-2/(4 + d)}.
+	\end{equation*}
 \end{enumerate}
@@
 
-Theorem~\ref{thm:laplacian_smoothing_testing1} is analogous to Theorem~\ref{thm:laplacian_eigenmaps_testing1}, except it deals with the Laplacian smoothing test statistic $T_{\LS}$ as opposed to $T_{\LE}$.
+Theorem~\ref{thm:laplacian_smoothing_testing1} is analogous to Theorem~\ref{thm:laplacian_eigenmaps_testing1}, except it deals with the Laplacian smoothing test $\wt{\phi}_{\LS}$ as opposed to $\wh{\phi}_{\LE}$.
 As in the estimation setting, we will require a tight range of scalings for $r$.
@@
 	\begin{equation*}
-	\biggl(\frac{\log n}{n}\biggr)^{1/d} \lesssim r(n) \lesssim M^{(16 - d)/(8 + 2d)}n^{(d - 20)/(32 + 8d)}
-	\end{equation*}
-	for all $n$ sufficiently large.
+	\biggl(\frac{\log n}{n}\biggr)^{1/d} \lesssim r(n) \lesssim M^{(16 - d)/(8 + 2d)}n^{(d - 20)/(32 + 8d)}.
+	\end{equation*}
 \end{enumerate}
@@
 	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d < 4$ and $f_0 \in H^1(\Xset,M)$ for $M \leq n^{(4 - d)/(4 + d)}$.
-Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}, and the Laplacian smoothing test $\wh{\phi}_{\LS}$ with $\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$, $s = 1$, and any $b \geq 1$.
+Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}, and the Laplacian smoothing test $\wt{\phi}_{\LS}$ with $\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$, $s = 1$, and any $b \geq 1$.
 Then for all $n$ sufficiently large, the following statement holds: if
@@
 \begin{itemize}
-	\item As mentioned in Section~\ref{subsec:sobolev_spaces}, when $d \geq 4$ we do not have minimax testing rates over $H^1(\Xset,M)$; these Sobolev balls include quite unsmooth functions, indeed functions without bounded $4$th moment, and it seems likely that no test---graph-based or otherwise--- can even be consistent against all alternatives in these classes.
+	\item As mentioned in Section~\ref{subsec:sobolev_spaces}, when $d \geq 4$ we do not have minimax testing rates over $H^1(\Xset,M)$; these Sobolev balls include irregular functions $f \not\in \Leb^4(\Xset)$, and it seems likely that no test---graph-based or otherwise--- can even be consistent against all alternatives in these classes.
 However, in Section~\ref{subsec:minimax_higher_order_smoothness} we derive upper bounds over those functions in $H^1(\Xset,M)$ which \emph{do} have bounded $4$th moment, in fact over all functions $f \in \Leb^4(\Xset)$.
 	\item To compute the threshold $\wt{t}_b$, one must know each of the eigenvalues $\lambda_1(G_{n,r}),\ldots,\lambda_n(G_{n,r})$.
-This is computationally more onerous than computing any of the estimators/statistics~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test}, which involve either finding only the first $\kappa$ eigenvalue/eigenvector pairs (Laplacian eigenmaps) or solving a linear system (Laplacian smoothing).
-That being said, this is not a serious impediment: in practice using asymptotically-optimal thresholds like $\wh{t}_b$ or $\wt{t}_b$ results in substantial loss of efficiency for moderate values of $n$, and we would make the standard recommendation to calibrate via permutation.
+Computing all $n$ of these eigenvalues is more expensive than computing any of the estimators/statistics~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test}, which involve either finding only the first $\kappa$ eigenvalue/eigenvector pairs (Laplacian eigenmaps) or solving a linear system (Laplacian smoothing).
+That being said, in practice we would not recommend using $\wt{t}_b$ or $\wh{t}_b$ anyway: the thresholds $\wh{t}_b$ and $\wt{t}_b$ are justified by asymptotic theory, but in practice $n$ is always finite, and using such thresholds can result in substantial loss of efficiency.
+We would instead make the standard recommendation to calibrate via permutation.
 \end{itemize}
 
-We now briefly set the stage for the rest of this Section.
-The estimation and testing rates established in Theorems~\ref{thm:laplacian_eigenmaps_estimation1}-\ref{thm:laplacian_smoothing_testing1} reflect a bonafide curse of dimensionality: as the dimension $d$ gets larger, inference quickly becomes (much) harder.
+We now set the stage for the rest of this Section.
+The estimation and testing rates established in Theorems~\ref{thm:laplacian_eigenmaps_estimation1}-\ref{thm:laplacian_smoothing_testing1} reflect a bonafide curse of dimensionality: as the dimension $d$ gets larger, inference quickly becomes much harder.
 As covered in Section~\ref{sec:problem_setup_and_background}, it is well known that by making additional structural assumptions---either on the smoothness of the regression function $f_0$, the dimension of the domain $\Xset$, or both---one can derive improved minimax rates.
-In the following two subsections, we show that under such assumptions, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} and test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} can in certain situations obtain the well-known faster minimax rates in the nonparametric estimation and testing problems.
+In the following two subsections, we show that under such assumptions, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} and test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} can in certain situations achieve the well-known faster minimax rates in the nonparametric estimation and testing problems.
 We treat these cases separately---from each other, and from the first-order full-dimensional case just covered---because we will require additional regularity conditions for the analysis to go through.
@@
 \label{subsec:minimax_estimation_higher_order}
-Now we would like to invoke Theorem~\ref{thm:graph_sobolev_seminorm}---as opposed to Lemma~\ref{lem:first_order_graph_sobolev_seminorm}---to upper bound the bias of $\wh{f}_{\LE}$: as a result, we will require that $r(n)$ satisfy the conditions of that Theorem.
+Now we would like to invoke Theorem~\ref{thm:graph_sobolev_seminorm}---as opposed to Lemma~\ref{lem:first_order_graph_sobolev_seminorm}---to upper bound the bias of $\wh{f}_{\LE}$: as a result, we will require that $r(n)$ be lower bounded as in the conditions of that Theorem.
 \begin{enumerate}[label=(R\arabic*)]
@@
 	\label{asmp:kernel_radius_higher_order_estimation}
-	\textbf{Neighborhood graph radius}:
 	The sequence of radii $r = r(n)$ satisfies
 	\begin{equation*}
-	n^{-1/(2(s - 1) + d)} \lesssim r(n) \lesssim M^{-2/(2s + d)}n^{-1/(2s + d)}
-	\end{equation*}
-	for all $n$ sufficiently large.
+	n^{-1/(2(s - 1) + d)} \lesssim r(n) \lesssim M^{-2/(2s + d)}n^{-1/(2s + d)}.
+	\end{equation*}
 \end{enumerate}
@@
 However, fundamentally speaking the assumption is still a mild one: we as the user are free to construct the neighborhood graph $G_{n,r}$ however we'd like.
-In Theorem~\ref{thm:laplacian_eigenmaps_estimation_higher_order}, we show that under such a scaling of $r$ and certain regularity conditions, the Laplacian eigenmaps estimator is, up to constants, asymptotically rate-optimal.
+In Theorem~\ref{thm:laplacian_eigenmaps_estimation_higher_order}, we show that under such a scaling of $r$ and certain regularity conditions, the Laplacian eigenmaps estimator is, up to constants, asymptotically optimal.
 
@@
 	\begin{equation*}
-	n^{-1/(2(s - 1) + d)} \leq r(n)~~\textrm{and}~~ r(n) \leq M^{-4/(4s + d)}n^{-2/(4s + d)}
-	\end{equation*}
-	for all $n$ sufficiently large.
+	n^{-1/(2(s - 1) + d)} \lesssim r(n)~~\textrm{and}~~ r(n) \lesssim M^{-4/(4s + d)}n^{-2/(4s + d)}.
+	\end{equation*}
 \end{enumerate}
 In one sense this is still a mild condition, in that the neighborhood graph radius is still a tuning parameter to be selected by the user.
-However, some elementary calculations show that when $d > 4$, it is in fact the case that $n^{-1/(2(s - 1) + d)} \gtrsim n^{-2/(4s + d)}$; hence no choice of $r$ can satisfy~\ref{asmp:kernel_radius_higher_order_testing}, unless $M$ is very small.
+However, some basic algebra shows that when $d > 4$, it is in fact the case that $n^{-1/(2(s - 1) + d)} \gg n^{-2/(4s + d)}$; hence no choice of $r$ can satisfy~\ref{asmp:kernel_radius_higher_order_testing}, unless $M$ is very small.
 As a result, our theory applies only when $d \leq 4$.
@@
 \begin{itemize}
-	\item \textbf{Low-smoothness regime}. \textcolor{red}{(TODO): Include results on minimax testing when $4s \leq d$.}
+	\item \textbf{Low-smoothness regime}. \textcolor{red}{(TODO 10)}: Include results on minimax testing when $4s \leq d$.
 	\item \textbf{Head off potential criticism}.
-\textcolor{red}{(TODO): Acknowledge frankly that there is a regime---when $4 < d < 4s$---where minimax testing rates are known, but for which we cannot prove that either of our test statistics achieve these rates.
-Our analysis will give some rates, but they will be suboptimal.}
+\textcolor{red}{(TODO 11)}: Acknowledge frankly that there is a regime---when $4 < d < 4s$---where minimax testing rates are known, but for which we cannot prove that either of our test statistics achieve these rates.
+Our analysis will give some rates, but they will be suboptimal.
 	\item \textbf{Whither Laplacian smoothing?} In the higher-order smoothness case, our analysis on the Laplacian smoothing methods is no longer tight enough, for all but a few values of $s$ and $d$, to imply minimax rate-optimality.
@@
 \label{sec:discussion}
-We have therefore shown that the properties of graph Sobolev classes established in Section~\ref{sec:graph_sobolev_classes} imply that graph-based procedures with high probability asymptotically achieve rate-optimal minimax risk, under appropriate regularity conditions.
+We have shown that the properties of graph Sobolev classes established in Section~\ref{sec:graph_sobolev_classes} imply that graph-based procedures with high probability asymptotically achieve rate-optimal minimax risk, under appropriate regularity conditions.
+
+\textcolor{red}{(TODO)}: Clean these up.
 \begin{itemize}
 	\item \textbf{Weaker conditions than required for ``continuum'' methods.}  We would like to emphasize one point regarding these regularity conditions: our analysis of graph-based procedures requires weaker assumptions on the design distribution $P$ than is needed to prove the equivalent results for their continuum counterparts.
-We require only that the density $p$ be bounded above and below, and that the domain $\Xset$ be Lipschitz (in the full-dimensional case) or smooth (in the manifold case), but crucially we do not assume that either the domain or density be known.
-This is in contrast with the continuum methods~\eqref{eqn:orthogonal_series} and~\eqref{eqn:smoothing_spline}, and it quantifies in a precise sense an advantage of graph-based methods over their continuum counterparts: the former automatically learn and leverage the shape of $p$ and $\Xset$ in a way the latter do not.
+On the density, we require only that $p$ be bounded above and below (the first-order case) or $C^{s-1}(\Xset)$ (the higher-order case); on the domain, that $\Xset$ be Lipschitz (in the full-dimensional case) or smooth (in the manifold case), but crucially we never assume that either the domain or density is known.
+This is in contrast with both orthogonal series and smoothing spline methods.
+For the former, we must always know $p$ and $\Xset$, in the sense that we always have access to an orthonormal basis of $\Leb^2(P)$.
+For the latter, under the manifold hypothesis one must know the manifold in order to compute $|\cdot|_{H^s(\Xset)}$.
+This formalizes an advantage of graph-based methods over their continuum counterparts: the former automatically learn and leverage the shape of $p$ and $\Xset$ in a way the latter do not.
+To be clear, we are not ruling out the possibility of tailoring these continuum methods to the case where $p$ and $\Xset$ are unknown; indeed some such tweaks have been proposed and analyzed, see for instance \textcolor{red}{(Mukherjee16)}.
+Our point is only that the graph-based procedures do not need any such tweaks---they work out of the box.
 	\item \textbf{Relationships between $s$ and $d$}.
@@
 	\item \textbf{Proof techniques}.
-On the subject of our proof techniques, the careful reader will note that our proofs do not rely on convergence of the Laplacian eigenvectors $v_k(G_{n,r})$ towards eigenfunctions of a limiting (weighted) Laplace-Beltrami operator.
-
-	\textcolor{red}{(TODO) Flesh the following out.} This is helpful for at least the following two reasons.
-First, proving convergence of eigenvectors is harder than proving convergence of eigenvalues, a fact which is reflected in the gap between the best known rates for the two problems.
-Second, because in order for our methods to be minimax optimal, orthogonal series estimators rely on projecting onto $\kappa$ basis functions, where $\kappa$ is a number growing in $n$.
-Thus we have an accumulation of error problem.
-	\item 
-		 In order to prove optimality of the continuum methods~\ref{eqn:orthogonal_series} and~\ref{eqn:smoothing_spline}, we require much stronger information on the design distribution $P$ than we do for our graph-based procedures.
-	\item \textbf{Possible extensions}. Methodologically: E.g. kNN graphs, different non-linear embedding schemes, etc. Theoretically: relaxing~\ref{asmp:bounded_density}, relaxing boundary conditions.
+On the subject of our proof techniques, the careful reader will note that our proofs do not rely on the convergence rates of the Laplacian eigenvectors $v_k(G_{n,r})$ towards eigenfunctions of a limiting (weighted) Laplace-Beltrami operator.
+This helps us avoid an accumulation of error problem.
+	\item \textbf{Possible extensions}.
+Methodologically: kNN graphs, different normalizations of the Laplacian, different graph-based procedures.
+Theoretically: relaxing~\ref{asmp:bounded_density}, relaxing boundary conditions.
 	\item \textbf{Boundary conditions}.
-Refer to \textcolor{red}{Belkin12} for a more precise description of the boundary conditions imposed by Laplacian eigenmaps, and \textcolor{red}{Ting18} for a catalog of potential solutions.
-Make clear that to ``fix'' the boundary bias problem---i.e.
-to obtain sharp rates for functions $f$ which do not satisfy $f \in H_0(\Xset)$, or at the least $f$ satisfying some Neumann boundary condition---likely requires a change of estimators, and is beyond the scope of this work.
+Refer to \textcolor{red}{(Belkin12)} for a more precise description of the boundary conditions imposed by Laplacian eigenmaps, and \textcolor{red}{(Ting18)} for a catalog of potential solutions.
+Reiterate that to ``fix'' the boundary bias problem---i.e.
+to obtain sharp rates for functions $f$ which do not satisfy $f \in H_0(\Xset)$, or at the least $f$ satisfying some Neumann boundary condition---we would likely have change of estimators, and is beyond the scope of this work.
 \end{itemize}
@@
 \begin{itemize}
-	\item \textbf{Estimation MSE 1.} Plot $L^2(P_n)$ error as a function of $n$ and choice $f_0^{(n)}$ when $P$ is (close to) uniform, showing that Laplacian methods are competitive with their non-graph-based brethren.
+	\item \textbf{Estimation error.} Set $P$ to be (close to) uniform. Plot error in empirical norm as a function of $n$ and regression functions $f_0^{(n)}$.
+	\item \textbf{Testing error.} Same thing but plotting power of our tests.
 	\item \textbf{Estimation fitted values.} Show fits for two regression problems, one where the distribution $P$ is assumed to be close to uniform, one where $P$ is dramatically non-uniform.
-	\item \textbf{Estimation MSE 2.} Plot $L^2(P_n)$ error as a function of the distribution $P$, as we vary $P$ from being uniform to being very non-uniform.
-
-	\item \textbf{Testing power.} 
+Show that in the former case the fitted values are similar, but in the latter case they are different.
+	\item \textbf{Manifold hypothesis.} Take a classic manifold---e.g. swiss roll---and do any of the above above.
+	\item \textbf{Comparing our graph-based methods.} For instance, show that when $d$ is large relative to $s$, the Laplacian smoothing method in fact does break down.
+Additionally, we could look at what happens for $s = 2$ or $s = 3$ to see which method in fact does a better job of taking advantage of higher-order smoothness, since our theory tells a less complete story in this case.
 \end{itemize}
-
-\begin{figure}
-	\centering
-	\includegraphics[width = .48\textwidth]{plots/graph_regression/mean_squared_errors/sobolev_MSE_by_density}
-	\caption{}
-	\label{fig:piecewise_cosine}
-\end{figure}
 

### 2020-09-16 notes/weekly_notes/8.26.20.NotesForWeek.tex

@@
 In this week's notes, I state and prove all results regarding graph Sobolev semi-norms that we need for our graph regression work.
+These notes make their way into Section 3 (theorem statements and remarks) and Appendix B (proofs) of our graph regression document.
 
@@
 \sum_{i = 1}^{n} \sum_{k \in [n]^q} \sum_{\ell \in [n]^q} \bigl(D_kf\bigr)(X_i) \cdot \bigl(D_\ell f\bigr)(X_i),~~ & \textrm{if $s$ is even, for $q = s/2$} \\
-\sum_{i,j = 1}^{n} \sum_{k \in [n]^q} \sum_{\ell \in [n]^q} \bigl(d_jD_kf\bigr)(X_i) \cdot \bigl(d_jD_{\ell}f\bigr)(X_i) \cdot K_r(X_i,X_j),~~ & \textrm{if $s$ is odd, for $q = (s - 1)/2$}
+\frac{1}{r^d}\sum_{i,j = 1}^{n} \sum_{k \in [n]^q} \sum_{\ell \in [n]^q} \bigl(d_jD_kf\bigr)(X_i) \cdot \bigl(d_jD_{\ell}f\bigr)(X_i) \cdot K_r(X_i,X_j),~~ & \textrm{if $s$ is odd, for $q = (s - 1)/2$}
 \end{dcases}
@@
 	\label{lem:expected_difference_lower_order_1}
-	Let $I = \abs{k \cup \ell \cup i} < 2q + 1$ for given index vectors $k,\ell \in [n]^q$ and index $i \in [n]$. Then for any $f \in H^1(\Xset)$,
+	Let $I = \abs{k \cup \ell \cup i} < 2q + 1 = s + 1$ for given index vectors $k,\ell \in [n]^q$ and index $i \in [n]$.
+If $f \in H^1(\Xset)$ for $\Xset$ a Lipschitz domain, and \textcolor{red}{$p$ satisfies (P2)}, then,
 	\begin{equation*}
-	\Bigl|\Ebb\Bigl[\bigl(D_kf\bigr)(X) \cdot \bigl(D_{\ell}f\bigr)(X)\Bigr]\Bigr| \leq C r^{2 + d(I - (s + 1))} \abs{f}_{H^1(\Xset)}^2
+	\Bigl|\Ebb\Bigl[\bigl(D_kf\bigr)(X_i) \cdot \bigl(D_{\ell}f\bigr)(X_i)\Bigr]\Bigr| \leq C r^{2 + d(I - (s + 1))} \abs{f}_{H^1(\Xset)}^2
+	\end{equation*}
+	for a constant $C$ which does not depend on $f$.
+\end{lemma}
+\begin{lemma}
+	\label{lem:expected_difference_lower_order_2}
+	Let $I = \abs{k \cup \ell \cup i \cup j} < 2q + 2 = s + 1$ for given index vectors $k,\ell \in [n]^q$ and indices $i,j \in [n]$.
+If $f \in H^1(\Xset)$ for $\Xset$ a Lipschitz domain, and \textcolor{red}{$p$ satisfies (P2)}, then,
+	\begin{equation*}
+	\Bigl|\Ebb\Bigl[\bigl(d_jD_kf\bigr)(X_i) \cdot \bigl(d_jD_{\ell}f\bigr)(X_i) \cdot K_r(X_i,X_j)\Bigr]\Bigr| \leq C r^{2 + d(I - s))} \abs{f}_{H^1(\Xset)}^2
 	\end{equation*}
@@
 	\end{equation}
-	and in Subsection~\ref{subsec:proof_sobolev_norm_boundary} we show that this inequality holds whenever $r > 0$ is sufficiently small.
+	and in \ref{subsec:proof_sobolev_norm_boundary} we show that this inequality holds whenever $r > 0$ is sufficiently small.
+\end{proof}
+
+\begin{proof}[\textbf{Proof of Lemma~\ref{lem:expected_difference_lower_order_1}}]
+	The proof of Lemma~\ref{lem:expected_difference_lower_order_2} will be essentially the same as that of Lemma~\ref{lem:expected_difference_lower_order_1}, and so we include only the proof of the latter.
+
+	We first reduce the problem of upper bounding the iterated difference operator to that of upper bounding a sum of single difference operators.
+Letting $k = (k_1,\ldots,k_q)$, we can show by induction that the absolute value of the iterated difference operator $\abs{D_kf(X_i)}$ is upper bounded,
+	\begin{equation*}
+	\abs{D_kf(X_i)} \leq \left(\frac{2K_{\max}}{r^d}\right)^{q-1} \sum_{h \in k \cup i} \abs{D_{k_q}f(X_h)} \cdot \1\Bigl\{G_{n,r}[X_{k \cup i}]~\textrm{is a connected graph} \Bigr\}.
+	\end{equation*}
+	Therefore, for $\ell = (\ell_1,\ldots,\ell_q)$, the product
+	\begin{align}
+	& \Bigl|D_kf(X_i)\Bigr| \cdot \Bigl|D_{\ell}f(X_i)\Bigr| \leq \nonumber \\
+	&~~ \left(\frac{2K_{\max}}{r^d}\right)^{2(q - 1)} \sum_{h,j \in k \cup \ell \cup i} \Bigl|D_{k_q}f(X_h)\Bigr| \cdot \Bigl|D_{\ell_q}f(X_j)\Bigr| \cdot \1\Bigl\{G_{n,r}[X_{k \cup i}], G_{n,r}[X_{\ell \cup i}]~\textrm{are connected graphs} \Bigr\} =  \nonumber \\
+	&~~ \left(\frac{2K_{\max}}{r^d}\right)^{2(q - 1)} \sum_{h,j \in k \cup \ell \cup i} \Big|D_{k_q}f(X_h)\Bigr| \cdot \Bigl|D_{\ell_q}f(X_j)\Bigr| \cdot \1\Bigl\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is a connected graph} \Bigr\} \label{pf:expected_difference_lower_order_1_0}
+	\end{align}
+
+	We now break our analysis into three cases, based on the number of distinct indices in $k_q,\ell_q,h,j$. In each case we will obtain the same rate
+	\begin{equation*}
+	\Ebb\Bigl[\abs{D_{k_q}f(X_h)} \cdot \abs{D_{\ell_q}f(X_j)}\Bigr] = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot |f|_{H^1(\Xset)}^2,
+	\end{equation*}
+	and plugging this back in to~\eqref{pf:expected_difference_lower_order_1_0} establishes the desired claim.
+
+	\textit{Case 1: Two distinct indices.}
+	Let $k_q = \ell_q = i$, and $h = j$. Using the law of iterated expectation, we obtain
+	\begin{align*}
+	\Ebb \left[\bigl(D_{i}f(X_j)\bigr)^2 \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\right] &= \Ebb \left[\bigl(D_{i}f(X_j)\bigr)^2 \cdot \Pbb\bigl[\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \} |X_i,X_j\bigr]\right] \\
+	& = O(r^{(\abs{k \cup \ell \cup i} - 2)d}) \cdot \Ebb\left[\bigl(D_{i}f(X_j)\bigr)^2\right] \\
+	& = O(r^{(\abs{k \cup \ell \cup i} - 3)d}) \cdot \Ebb\left[\bigl(d_{i}f(X_j)\bigr)^2K_r(X_i,X_j)\right] \\
+	& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot |f|_{H^1(\Xset)}^2
+	\end{align*}
+	where we established the last equality in the proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm}.
+
+	\textit{Case 2: Three distinct indices.}
+	Let $k_q = \ell_q = i$, for some $i \neq j \neq h$. Using the law of iterated expectation, we obtain
+	\begin{align}
+	\Ebb \Bigl[ & \abs{D_{i}f(X_j)} \cdot \abs{D_if(X_h)} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\Bigr] \nonumber \\
+	& = \Ebb\Bigl[\abs{D_{i}f(X_j)} \cdot \abs{D_if(X_h)} \cdot \Pbb\bigl[\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \} |X_i,X_j,X_h\bigr]\Bigr] \nonumber \\
+	& = O(r^{(\abs{k \cup \ell \cup i} - 3)d}) \cdot \Ebb\Bigl[\abs{D_{i}f(X_j)} \cdot \abs{D_if(X_h)} \Bigr]\nonumber \\
+	& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2})\cdot|f|_{H^1(\Xset)}^2. \label{pf:expected_difference_lower_order_1_1}
+	\end{align}
+	where we prove~\eqref{pf:expected_difference_lower_order_1_1} in~\ref{subsec:proof_expected_difference_lower_order}.
+
+	\textit{Case 3: Four distinct indices.}
+	Using the law of iterated expectation, we find that
+	\begin{align}
+	\Ebb\Bigl[ &\abs{D_{k_q}f(X_i)}\cdot{\abs{D_{\ell_q}f(X_j)}} \cdot \1\{G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected} \}\Bigr] \nonumber \\
+	& = \Ebb\Bigl[\abs{D_{k_q}f(X_i)}\cdot{\abs{D_{\ell_q}f(X_j)}} \cdot\Pbb\bigl[G_{n,r}[X_{k \cup \ell \cup i}] ~\textrm{is connected}|X_i,X_j,X_{k_q},X_{\ell_q}\bigr]\Bigr] \nonumber \\
+	& = O(r^{(\abs{k \cup \ell \cup i} - 4)d}) \cdot \Ebb\Bigl[\abs{D_{k_q}f(X_i)}\cdot{\abs{D_{\ell_q}f(X_j)}}\cdot\1\{\norm{X_i - X_j} \leq (2q + 1)r\}\Bigr] \nonumber \\
+	& = O(r^{(\abs{k \cup \ell \cup i} - 3)d + 2}) \cdot |f|_{H^1(\Xset)}^2 \label{pf:expected_difference_lower_order_1_2}
+	\end{align}
+	where we prove~\eqref{pf:expected_difference_lower_order_1_2} in~\ref{subsec:proof_expected_difference_lower_order}.
 \end{proof}
@@
 \begin{proof}[\textbf{Proof of Theorem~\ref{thm:graph_sobolev_seminorm}}]
-	From the representation~\eqref{eqn:graph_sobolev_seminorm_1}, we have that 
+	We will divide our proof into two cases: when $s$ is even and when $s$ is odd.
+
+	\textit{Case 1: $s$ is even.} We gather from~\eqref{eqn:graph_sobolev_seminorm_1} that 
 	\begin{equation}
@@
 	with the second inequality following from~\eqref{eqn:norm_expected_difference_operator}.
-There are at most $O(n^{s + 1})$ such terms, hence the overall contribution of these terms to the expectation~\eqref{pf:graph_sobolev_seminorm} is $O(n^{s + 1}r^{s(d + 2)})M^2$.
+Put simply, for those terms in~\eqref{pf:graph_sobolev_seminorm} such that $\abs{k \cup \ell \cup i} = 2q + 1$ the expectation $\Ebb\bigl[\bigl(D_kf\bigr)(X_i) \cdot \bigl(D_{\ell}f\bigr)(X_i)\bigr] = O(r^{s(d + 2)})M^2$.
+There are at most $O(n^{s + 1})$ such terms, hence the overall contribution of these terms to the sum in~\eqref{pf:graph_sobolev_seminorm} is $O(n^{s + 1}r^{s(d + 2)})M^2$.
 
@@
 Taking $I = s$ to be the largest integer less than $s + 1$, the contribution of these terms to the sum is therefore $O(n^sr^{d(s - 1) + 2})M^2$, which in light of the restriction $r \geq n^{-1/(2(s - 1) + d)}$ is $O(n^{s+1}r^{s(d +2)})M^2$.
+
+	\textit{Case $2$: $s$ is odd.} We see from~\eqref{eqn:graph_sobolev_seminorm_1} that for $q = (s - 1)/2$,
+	\begin{equation}
+	\label{pf:graph_sobolev_seminorm_4}
+	\Ebb\Bigl[f^T \Lap_{n,r}^s f\Bigr] = r^{d(s - 1)}\sum_{i,j = 1}^{n} \sum_{k \in [n]^q} \sum_{\ell \in [n]^q} \Ebb\Bigl[\bigl(d_jD_kf\bigr)(X_i) \cdot  \bigl(d_jD_{\ell}f\bigr)(X_i) \cdot K_r(X_i,X_j)\Bigr]
+	\end{equation}
+	and we divide into cases based on the cardinality of $\abs{k \cup \ell \cup i \cup j}$.
+If $\abs{k \cup \ell \cup i \cup j} = 2q + 2$, the intersection between any two of $k \in (n)^q$, $\ell \in (n)^q$, $i \in [n]$ and $j \in [n]$ is empty, and the law of iterated expectation implies
+	\begin{align}
+	\Ebb\Bigl[\bigl(d_jD_kf\bigr)(X_i) \cdot  \bigl(d_jD_{\ell}f\bigr)(X_i) \cdot K_r(X_i,X_j)\Bigr] & =\Ebb\Bigl[\Bigl\{\bigl(d_j\Delta_r^qf\bigr)(X_i)\Bigr\}^2  K_r(X_i,X_j)\Bigr] \nonumber \\
+	&\leq p_{\max}^2 \int_{\Xset} \int_{\Xset} \Bigl(\Delta_r^qf(x') - \Delta_r^qf(x)\Bigr)^2 K_r(x',x) \,dx' \,dx \label{pf:graph_sobolev_seminorm_3}
+	\end{align}
+	From here the basic idea is the same as in the case where $s$ was even. We partition the domain of integration into $\Xset = \Xset_{qr} \cup \Xset^{qr}$, with the result being that
+	\begin{align}
+	\int_{\Xset} & \int_{\Xset} \Bigl(\Delta_r^qf(x') - \Delta_r^qf(x)\Bigr)^2 K_r(x',x) \,dx' \,dx \nonumber \\ & = \int_{\Xset} \biggl\{\int_{\Xset_{qr}} \Bigl(\Delta_r^qf(x') - \Delta_r^qf(x)\Bigr)^2 K_r(x',x) \,dx' + \int_{\Xset^{qr}} \Bigl(\Delta_r^qf(x') - \Delta_r^qf(x)\Bigr)^2 K_r(x',x) \,dx'\biggr\}\,dx \nonumber \\
+	& \leq \int_{\Xset_{qr}} \int_{\Xset_{qr}} \Bigl(\Delta_r^qf(x') - \Delta_r^qf(x)\Bigr)^2 K_r(x',x) \,dx' \,dx + 2 \int_{\Xset} \int_{\Xset^{qr}} \Bigl(\Delta_r^qf(x') - \Delta_r^qf(x)\Bigr)^2 K_r(x',x) \,dx'\,dx \nonumber \\
+	& \leq \int_{\Xset_{qr}} \int_{\Xset_{qr}} \Bigl(\Delta_r^qf(x') - \Delta_r^qf(x)\Bigr)^2 K_r(x',x) \,dx' \,dx + 8 r^d \int_{\Xset^{(q + 1)r}} \Bigl(\Delta_r^qf(x)\Bigr)^2 \,dx \label{pf:graph_sobolev_seminorm_1}
+	\end{align}
+	with the inequality~\eqref{pf:graph_sobolev_seminorm_1} following since the kernel $K$ is supported on $[0,1]$ and integrates to $1$.
+
+	Now, noticing that the second integral in~\eqref{pf:graph_sobolev_seminorm_1} is exactly $\norm{\Delta_r^{q}f}_{\Leb^2(\Xset^{(q + 1)r})}$, we have from Lemma~\ref{lem:expected_difference_boundary} that
+	\begin{equation}
+	\label{pf:graph_sobolev_seminorm_5}
+	\int_{\Xset^{(q + 1)r}} \Bigl(\Delta_r^qf(x)\Bigr)^2 \leq CMr^{2s}.
+	\end{equation}
+	For the first integral in~\eqref{pf:graph_sobolev_seminorm_1}, we apply the triangle inequality and obtain
+	\begin{align}
+	\int_{\Xset_{qr}} & \int_{\Xset_{qr}} \Bigl(\Delta_r^qf(x') - \Delta_r^qf(x)\Bigr)^2 K_r(x',x) \,dx' \,dx \nonumber \\
+	& \leq 6 \int_{\Xset_{qr}} \int_{\Xset_{qr}} \Bigl(\Delta_r^qf(x) - \Delta_p^qf(x) \Bigr)^2 K_r(x',x) \,dx' \,dx + 3 \int_{\Xset_{qr}} \int_{\Xset_{qr}} \Bigl(\Delta_p^qf(x') - \Delta_p^qf(x) \Bigr)^2 K_r(x',x) \,dx' \,dx \nonumber \\
+	& \leq 6r^d \int_{\Xset_{qr}} \int_{\Xset_{qr}} \Bigl(\Delta_r^qf(x) - \Delta_p^qf(x) \Bigr)^2 \,dx + 3 \int_{\Xset_{qr}} \int_{\Xset_{qr}} \Bigl(\Delta_p^qf(x') - \Delta_p^qf(x) \Bigr)^2 K_r(x',x) \,dx' \,dx \nonumber \\
+	& \leq c M r^{2s + d} \label{pf:graph_sobolev_seminorm_2}
+	\end{align}
+	where the last inequality follows from Lemma~\ref{lem:expected_difference_sobolev}---which bounds the first term---and the proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm}---which bounds the second term since $\Delta_p^qf \in H^1(\Xset)$ for $q = (s - 1)/2$.
+Combining~\eqref{pf:graph_sobolev_seminorm_3}-\eqref{pf:graph_sobolev_seminorm_2}, we have that for those terms inside the sum~\eqref{pf:graph_sobolev_seminorm_4} with a total of $\abs{k \cup \ell \cup i \cup j} = s$ distinct indices, the expectation $\Ebb\bigl[(d_jD_kf)(X_i) \cdot (d_jD_{\ell}f)(X_i)\bigr] = O(r^{s(d + 2)})M^2$.
+
+	Otherwise $\abs{k \cup \ell \cup j \cup i} = I < 2q + 2$.
+By Lemma~\ref{lem:expected_difference_lower_order_2} the total contribution of such terms inside the sum~\eqref{pf:graph_sobolev_seminorm} is $O(n^{I}r^{d(I - 1) + 2}) M^2$, which in turn is $O(n^{s + 1}r^{s(d + 2)})M^2$ by exactly the same used in the first case, when $s$ was even.
 \end{proof}
@@
 
+\section{Integrals}
+We compute several bounds on several integrals referenced in various proofs.
+
 \subsection{Proof of~\eqref{eqn:integrated_remainder_term_sobolev} and~\eqref{eqn:integrated_remainder_term_holder}}
@@
 
+\subsection{Proof of~\eqref{pf:expected_difference_lower_order_1_1} and~\eqref{pf:expected_difference_lower_order_1_2}}
+\label{subsec:proof_expected_difference_lower_order}
+
+We will assume that $f \in C^{\infty}(\Xset)$, without loss of generality due to the density of $C^{\infty}(\Xset)$ in $H^1(\Xset)$.
+Additionally, since $\Xset$ is a Lipschitz domain, we may take $g \in C^{\infty}(\Rd)$ to be an extension of $f$ such that $g = f$ almost everywhere on $\Xset$, and
+\begin{equation*}
+|g|_{H^1(\Rd)} \leq C|f|_{H^1(\Xset)},
+\end{equation*}
+Since $g = f$ almost everywhere on $\Xset$, it suffices to establish the inequalities~\eqref{pf:expected_difference_lower_order_1_1} and~\eqref{pf:expected_difference_lower_order_1_2} hold with respect to $g$.
+
+We begin by proving~\eqref{pf:expected_difference_lower_order_1_1}, following a very similar series of steps to the proof of~\eqref{eqn:norm_expected_difference_operator} when $s = 1$.
+First, we account for the effect of the density $p$ using $|p(x)| \leq p_{\max}$,
+\begin{align}
+\Ebb\Bigl[\abs{D_{i}g(X_j)} \cdot \abs{D_ig(X_h)} \Bigr] & = \frac{1}{r^{2d}}\int_{\Xset} \int_{\Xset} \int_{\Xset} \bigl|g(x'') - g(x)\bigr| \cdot \bigl|g(x') - g(y)\bigr| K_r(x'',x) K_r(x',x) p(x'') p(x') p(x) \,dx'' \,dx' \,dx \nonumber \\
+& = \int \left[\frac{1}{r^d}\int_{\Xset} \bigl|g(x') - g(x)\bigr| K_r(x',x) p(x') \,dx'\right]^2 p(x) \,dx \nonumber \\
+& \leq p_{\max}^3 \left[\frac{1}{r^d}\int_{\Xset} \bigl|g(x') - g(x)\bigr| K_r(x',x) \,dx'\right]^2 \,dx \label{pf:proof_expected_difference_lower_order_1}
+\end{align}
+Then taking a Taylor expansion of $g$ about $x' = x$ gives 
+\begin{align*}
+\int_{\Xset} \bigl|g(x') - g(x)\bigr| K_r(x',x) \,dx' & \leq \int_{\Rd} \bigl|g(x') - g(x)\bigr| K_r(x',x) \,dx' \\
+& = \int_{\Rd} \abs{\int_{0}^{1} (1 - t) \sum_{i = 1}^{d} g^{(e_i)}(x + t(x' - x)) (x' - x)^{e_i} \,dt} K_r(x',x) \,dx \\
+& \leq \int_{\Rd} \int_{0}^{1} \sqrt{\sum_{i = 1}^{d}\Bigl(g^{(e_i)}\bigl(x + t(x' - x)\bigr)\Bigr)^2} \cdot \bigl\|x - x'\bigr\| \,dt K_r(x',x) \,dx \\
+& \leq r \int_{\Rd} \int_{0}^{1} \sqrt{\sum_{i = 1}^{d}\Bigl(g^{(e_i)}\bigl(x + t(x' - x)\bigr)\Bigr)^2} \,dt K_r(x',x) \,dx \\
+& \leq K_{\max} r^{d + 1} \int_{B(0,1)} \int_{0}^{1} \sqrt{\sum_{i = 1}^{d}\Bigl(g^{(e_i)}(x + tzr)\Bigr)^2} \,dt \,dz,
+\end{align*}
+whence the claim~\eqref{pf:expected_difference_lower_order_1_1} follows from plugging back in to~\eqref{pf:proof_expected_difference_lower_order_1}, then using Jensen's inequality and Fubini's inequality as in~\ref{subsec:proof_integrated_remainder_terms}.
+
+We prove~\eqref{pf:expected_difference_lower_order_1_2} in similar fashion. First we account for the density,
+\begin{align*}
+\Ebb\Bigl[& \bigl|D_{k}g(X_i)\bigr|\cdot \bigl|D_{\ell}g(X_j)\bigr| \cdot\1\bigl\{\norm{X_i - X_j} \leq (2q + 1)r\bigr\}\Bigr] \\
+& \leq \frac{p_{\max}^4}{r^{2d}} \int_{\Xset^4} \bigl|g(x''') - g(x'')\bigr| \cdot \bigl|g(x') - g(x)\bigr| \cdot  K_r(x''',x'') K_r(x',x) \1\bigl\{\norm{x''' - x} \leq (2q + 1)r\bigr\} \,dx''' \,dx'' \,dx' \,dx.
+\end{align*}
+Substituting $z = (x''' - x)/r$, $z' = (x' - x)/r$, and $z'' = (x''' - x'')/r$, we can simplify the integral in the previous display,
+\begin{align*}
+\int_{\Xset^4} & \bigl|g(x''') - g(x'')\bigr| \cdot \bigl|g(x') - g(x)\bigr| \cdot  K_r(x''',x'') K_r(x',x) \1\bigl\{\norm{x''' - x} \leq (2q + 1)r\bigr\} \,dx''' \,dx'' \,dx' \,dx \\
+& \leq K_{\max}^2 r^{3d} \int_{\Xset} \int_{B(0,2q + 1)^3} \big|g\bigl((x + r(z + z'')\bigr) - g(x + rz)\bigr| \cdot \bigl|g(x + z'r) - g(x)\bigr| \,dz'' \,dz' \,dz \,dx.
+\end{align*}
+From here the claim~\eqref{pf:expected_difference_lower_order_1_2} follows upon Taylor expansion of $g$, in essentially the same fashion as the proof of~\eqref{pf:expected_difference_lower_order_1_1}.
+
 \subsection{Proof of~\eqref{pf:expected_difference_boundary}}

### 2020-09-28 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing.tex
First draft---in outline form---of AISTATS paper on Laplacian smoothing.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{tikz}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\iid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}
\newcommand{\Lap}{L}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}
\newcommand{\Id}{I}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Vset}{\mathcal{V}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Minimax optimal Laplacian smoothing}
\author{Alden Green}
\date{\today}
\maketitle

\textbf{Disclaimer:} This largely follows the same setup as in the draft I have written which combines Laplacian smoothing and Laplacian eigenmaps results. That draft --- document name \textit{graph\_regression2} --- should for the moment be considered obsolete.

\section{Introduction}

\textbf{(1)} Laplacian smoothing is a graph-based approach to nonparametric regression.

\begin{itemize}
	\item In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Rd$, and 
	\begin{equation}
	\label{eqn:random_design_regression}
	Y_i = f_0(X_i) + \varepsilon_i
	\end{equation}
	with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise. Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present. 
	\item In graph-based nonparametric regression, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$. The neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$. The $n \times n$ weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $\mathbf{W}_{ij}$ are given by
	\begin{equation*}
	\label{eqn:neighborhood_graph}
	{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr).
	\end{equation*}
	Then the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$, and the graph Laplacian can be written as
	\begin{equation}
	\label{eqn:graph_Laplacian}
	\Lap_{n,r} = \bf{D} - \bf{W}
	\end{equation}
	\item The Laplacian smoothing estimator $\wh{f}_{\LS}$ \citep{smola2003} is a penalized least squares estimator, given by
	\begin{equation}
	\label{eqn:laplacian_smoothing}
	\wh{f}_{\LS} := \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_{n,r} f \biggr\}
	\end{equation}
	where $\rho > 0$ acts a tuning parameter on the penalty $f^T \Lap_{{n,r}} f$.
	Assuming~\eqref{eqn:laplacian_smoothing} is a reasonable estimator of $f_0$, the squared empirical norm
	\begin{equation}
	T_{\LS} = \frac{1}{n}\Bigl\|\wt{f}_{\LS}\Bigr\|_2^2 \label{eqn:laplacian_smoothing_test}
	\end{equation}
	is in turn a reasonable statistic to assess whether or not $f_0 = 0$. 
\end{itemize}

\textbf{(2.a)} Laplacian smoothing has many advantages compared to other nonparametric regression methods. For instance:

\begin{itemize}
	\item It is fast, easy, and stable to compute. Letting ${\bf Y} = (Y_1,\ldots,Y_n) \in \Reals^n$, it is not hard to see that the minimizer $\wt{f}_{\LS} = (\rho \Lap_{n,r} + I)^{-1}{\bf Y}$ can be computed by solving a linear equation; in practice, one typically chooses $r$ to be small, with the result being that the matrix $\Lap_{n,r}$ is sparse. There exist known fast solvers of exactly this system.
	\item Is is generalizable to non-standard data---for instance, text or images, or really any data modality on which it is possible to define a kernel.
	\item It is easily adaptable to the semi-supervised learning setting. 
\end{itemize}

\textbf{(2.b)} For this reason a considerable body of work has emerged analyzing the consistency of the graph Laplacian $\Lap_{{n,r}}$ as $n \to \infty$ and $r(n) \to 0$. This is meant in various senses:
\begin{itemize}
	\item Pointwise consistency, meaning $\Lap_{n,r}f \to \Delta_Pf$ for an appropriate limiting operator $\Delta_P$.
	\item Spectral consistency, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$.
	\item Consistency of norms, meaning $f^T \Lap_{n,f} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$. 
\end{itemize}
However, the statistical optimality of Laplacian smoothing is still not well understood.

\textbf{(2.c)} Our main contributions lie in this latter direction, and they can all be summarized as follows: Laplacian smoothing methods are minimax optimal over Sobolev spaces. In more detail, we will show that when $f_0$ belongs to the Sobolev ball $H^1(\Xset;M)$:
\begin{itemize}
	\item \textbf{Minimax optimal estimation.} With high probability, $\frac{1}{n}\norm{\wt{f}_{\LS} - f_0}_{2}^2 \lesssim n^{-2/(2 + d)}$.
	\item \textbf{Minimax optimal testing.}
	A level-$\alpha$ test constructed using $T_{\LS}$ has non-trivial power whenever $\norm{f_0}_{\Leb^2(\Xset)}^2 \gtrsim n^{-4/(4 + d)}$. 
	\item \textbf{Manifold adaptivity.}
	If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$, both of the aforementioned rates hold with $d$ replaced by $m$.
\end{itemize}
Finally, we will also establish that the Laplacian smoothing estimator can be altered to take advantage of higher-order smoothness assumptions on $f_0$. This last conclusion will hold only under very limited circumstances.

\paragraph{Organization.}

List out what is to come.

\section{Minimax regression over Sobolev spaces}

\textbf{(3)} Sobolev spaces are classical function spaces, over which linear nonparametric regression methods. One notable such method is smoothing splines, to which we give special attention because of their close connection to Laplacian smoothing.

\textbf{(4)} In fact, an alternative way to motivate Laplacian smoothing nonparametric regression methods---as opposed to the advantages mentioned in \textbf{(2.a)}---comes from viewing them as discrete-time versions of smoothing splines. This is made more rigorous by the consistency results of \textbf{(2.b)}. Viewed in this light, an additional advantage of Laplacian smoothing emerges: it automatically adapts to the geometry of $\Xset$ and $p$.

\textbf{(5)} Smoothing splines are known to have strong minimax optimal properties. Reasoning by analogy, we might hope that Laplacian smoothing would share these properties.

\section{Graph Sobolev classes}

\textbf{(6)} The minimax optimality of smoothing splines rests on two crucial facts: one, the a priori assumption that $|f|_{H^1(\Xset)}$ is bounded; two, an upper bound on the metric entropy of $H^1(\Xset)$, characterized by the decay of its eigenvalues. \textcolor{red}{(TODO)}:(I know this isn't well-stated, and I need to clean it up.)  We now show that qualitatively similar statements hold with respect to

\quad \textbf{(6.1)} the graph Sobolev semi-norm $f_0^T \Lap_{n,r} f_0$ and,

\quad \textbf{(6.2)} the graph eigenvalues $\lambda_k(\Lap_{n,r})$.

\section{Minimax optimality of Laplacian smoothing}

\textbf{(7)} Using \textbf{(6.1)} and \textbf{(6.2)}, we establish upper bounds on the estimation and testing error of Laplacian smoothing, which confirm that Laplacian smoothing is asymptotically minimax rate-optimal over the Sobolev class $H^1(\Xset)$.

\textbf{(8)} While we motivated graph Laplacians by suggesting they might replicate the statistical properties of smoothing splines, we have in fact managed to show something stronger. In particular:

\quad \textbf{(8.1)} When $d \geq 2$, there is no Sobolev embedding of $H^1(\Xset) \subseteq C^s(\Xset)$ for any $s > 0$. Therefore the smoothing spline estimator is ill-posed.

\quad \textbf{(8.2)} Additionally, the metric entropy of $C^1(\Xset)$ is too large to obtain optimal rates for \textcolor{red}{Holder-smoothing splines} when $d \geq 2$. Thus, we see a gap emerge between the optimality properties of the discrete-time (Laplacian smoothing) and continuous-time (smoothing splines) estimators, in favor of the latter.

\quad \textbf{(8.3)} On the other hand, when $d \geq 4$, we no longer get the known optimal rates in the estimation problem. When $d = 4$, our rates are suboptimal by a factor of $\log n$. When $d > 5$ they are very suboptimal. We believe this is related to classical results regarding the entropy of low-smoothness Holder classes, i.e. the results used to justify \textbf{(8.2)}. The interesting point is that the parameter spaces $C^1(\Xset)$ and $H^1(G_{n,r})$ become ``too large'' for different values of $d$.

\quad \textbf{(8.4)} \textcolor{red}{(TODO)}: Translate these results into error in $\Leb^2(P)$ error, if possible.

\subsection{Manifold adaptivity}

\quad \textbf{(9)} We now show that when $\Xset$ is an $m$-dimensional submanifold of $\Reals^d$, Laplacian smoothing approaches automatically ``feel'' the intrinsic dimension $m$ of $\Xset$, which translate into improved upper bounds on estimating and testing error. This confirms the statement in \textbf{(4)} that Laplacian smoothing automatically adapts to the geometry of $\Xset$ and $p$.

\subsection{Higher-order smoothness classes}

\quad \textbf{(10)} Analogously to smoothing splines, Laplacian smoothing can also be adapted to take advantage of additional assumed regularity on $f_0$, i.e. $f_0 \in H^s(\mc{\Xset})$ for $s > 1$. In very limited circumstances, we can show that this adapted estimator achieves the sharper minimax rates over these classes. However, the general story--for all combinations of $s$ and $d$---remains beyond our reach.

\section{Simulations}

\quad \textbf{(11)} Empirically, we demonstrate that the risk for Laplacian smoothing is comparable to that of smoothing splines, in the full dimensional case. Although one cannot carry out a comparison in the case of an unknown manifold --- since the smoothing spline estimator must be changed in a non-obvious manner --- we do show that when the density $p$ is concentrated close to a manifold and $f_0$ is smooth where $p$ is peaked, we have improved performance for $\wh{f}_{\LS}$ compared to $\wh{f}_{\mathrm{SM}}$.

\section{Discussion}

\clearpage

\bibliographystyle{plainnat}
\bibliography{../../../graph_testing_bibliography}

\end{document}

### 2020-09-30 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing2.tex
Draft of AIStats submission added, with sections 1-3 now filled in.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{tikz}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\iid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}
\newcommand{\Lap}{L}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}
\newcommand{\Id}{I}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Vset}{\mathcal{V}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\SM}{\mathrm{SM}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Minimax optimal Laplacian smoothing}
\author{Alden Green}
\date{\today}
\maketitle

\textbf{Disclaimer:} This largely follows the same setup as in the draft I have written which combines Laplacian smoothing and Laplacian eigenmaps results. That draft --- document name \textit{graph\_regression2} --- should for the moment be considered obsolete.

\section{Introduction}

\textbf{(1)} Laplacian smoothing is a graph-based approach to nonparametric regression.

\begin{itemize}
	\item In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Reals^d$, and 
	\begin{equation}
	\label{eqn:random_design_regression}
	Y_i = f_0(X_i) + \varepsilon_i
	\end{equation}
	with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise. We will assume throughout that $P$ admits a density $p$ with respect to the volume form of $\mc{X}$ which is bounded away from $0$ and $\infty$, i.e.
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty,~~\textrm{for all $x \in \Xset$.}
	\end{equation*}
	Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present. 
	\item In graph-based nonparametric regression, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$. The neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$. The $n \times n$ weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $\mathbf{W}_{ij}$ are given by
	\begin{equation*}
	\label{eqn:neighborhood_graph}
	{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr).
	\end{equation*}
	Then the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$, and the graph Laplacian can be written as
	\begin{equation}
	\label{eqn:graph_Laplacian}
	\Lap_{n,r} = \bf{D} - \bf{W}
	\end{equation}
	\item The Laplacian smoothing estimator $\wh{f}_{\LS}$ \citep{smola2003} is a penalized least squares estimator, given by
	\begin{equation}
	\label{eqn:laplacian_smoothing}
	\wh{f}_{\LS} := \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_{n,r} f \biggr\}
	\end{equation}
	where $\rho > 0$ acts a tuning parameter on the penalty $f^T \Lap_{{n,r}} f$.
	Assuming~\eqref{eqn:laplacian_smoothing} is a reasonable estimator of $f_0$, the squared empirical norm
	\begin{equation}
	T_{\LS} = \frac{1}{n}\Bigl\|\wt{f}_{\LS}\Bigr\|_2^2 \label{eqn:laplacian_smoothing_test}
	\end{equation}
	is in turn a reasonable statistic to assess whether or not $f_0 = 0$. 
\end{itemize}

\textbf{(1.a) Motivation.} Laplacian smoothing has many advantages compared to other nonparametric regression methods. For instance:

\begin{itemize}
	\item It is fast, easy, and stable to compute. Letting ${\bf Y} = (Y_1,\ldots,Y_n) \in \Reals^n$, it is not hard to see that the minimizer $\wt{f}_{\LS} = (\rho \Lap_{n,r} + I)^{-1}{\bf Y}$ can be computed by solving a linear equation; in practice, one typically chooses $r$ to be small, with the result being that the matrix $\Lap_{n,r}$ is sparse. There exist known fast solvers of exactly this system.
	\item Is is generalizable to non-standard data---for instance, text or images, or really any data modality on which it is possible to define a kernel.
	\item It is easily adaptable to the semi-supervised learning setting. 
\end{itemize}

\textbf{(1.b) Related work.} For this reason a considerable body of work has emerged analyzing the convergence of the graph Laplacian $\Lap_{{n,r}}$ to the continuum Laplace operator $\Delta_P: C^2(\Xset) \to \Leb^2(\Xset)$,
\begin{equation}
\label{eqn:laplace_operator}
\Delta_Pg := -\frac{1}{p} \dive(p^2 \nabla g)
\end{equation}
as $n \to \infty$ and $r(n) \to 0$. This is meant in various senses:
\begin{itemize}
	\item Pointwise convergence, meaning $\Lap_{n,r}f \to \Delta_Pf$. 
	\item Spectral convergence, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$.
	\item Convergence of norms, meaning $f^T \Lap_{n,r} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$. 
\end{itemize}
However, the statistical optimality of Laplacian smoothing is still not well understood.

\textbf{(1.c) Our contributions.} Our main contributions lie in this latter direction, and they can all be summarized as follows: Laplacian smoothing methods are minimax optimal over Sobolev spaces. In more detail, we will show that when $f_0$ belongs to the Sobolev ball $H^1(\Xset;M)$:
\begin{itemize}
	\item \textbf{Minimax optimal estimation.} With high probability, $\frac{1}{n}\norm{\wt{f}_{\LS} - f_0}_{2}^2 \lesssim n^{-2/(2 + d)}$.
	\item \textbf{Minimax optimal testing.}
	A level-$\alpha$ test constructed using $T_{\LS}$ has non-trivial power whenever $\norm{f_0}_{\Leb^2(\Xset)}^2 \gtrsim n^{-4/(4 + d)}$. 
	\item \textbf{Manifold adaptivity.}
	If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$, both of the aforementioned rates hold with $d$ replaced by $m$.
\end{itemize}
Finally, we will also establish that the Laplacian smoothing estimator can be altered to take advantage of higher-order smoothness assumptions on $f_0$. This last conclusion will hold only under very limited circumstances.

\textbf{(1.d) Organization.}

\textcolor{red}{(TODO)}

\textbf{(1.e) Notation.}

\textcolor{red}{(TODO)}

\section{Minimax-optimal regression over Sobolev spaces}

\textbf{(2)} Before we get to our main results, we briefly review minimax optimal estimation and testing rates over Sobolev classes. We also single out a notable method which achieves these rates: \emph{smoothing splines}. As we will see, Laplacian smoothing can be seen as a graph-based---i.e. discrete---counterpart to the ``continuous-time'' approach of smoothing splines.

\paragraph{(2.a) Sobolev spaces.} We start by briefly introducing Sobolev spaces, emphasizing only the material relevant to our statistical context. 
\begin{itemize}
	\item Roughly speaking, Sobolev spaces contain functions $f \in \Leb^p(\Xset)$ with derivatives which themselves belong to $\Leb^p(\Xset)$. More formally, for given integers $s$ and $p > 0$, the Sobolev space $W^{s,p}(\Xset)$ consists of all functions $f \in \Leb^p(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ exists and belongs to $\Leb^p(\Xset)$. For such functions, the $(s,p)$-Sobolev seminorm $\seminorm{f}_{W^{s,p}(\Xset)}$ and norm $\norm{f}_{W^{s,p}(\Xset)}$ are given by 
	\begin{equation*}
	\seminorm{f}_{W^{s,p}(\Xset)}^p = \sum_{\abs{\alpha} = s}\int_{\mathcal{X}} \Bigl|\bigl(D^{\alpha}f\bigr)(x)\Bigr|^p \,dx, ~~ \norm{f}_{W^{s,p}(\Xset)}^p = \sum_{k = 0}^{s} \seminorm{f}_{W^{k,p}(\Xset)}^p
	\end{equation*}
	and for a given $M > 0$, the $(s,p)$-Sobolev ball is $W^{s,p}(\Xset, M) = \set{f: \norm{f}_{W^{s,p}(\Xset)} \leq M}$. In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space, and we adopt the usual convention of writing $H^s(\Xset) = W^{s,2}(\Xset)$ and $H^s(\Xset,M) = W^{s,2}(\Xset,M)$. Linear estimators such as $\wh{f}_{\LS}$\footnote{To be explicit, estimators of the form $\wh{f} = HY$ for some matrix $H \in \Reals^{n \times n}$.} can be very accurate over the Hilbert-Sobolev spaces $H^s(\Xset)$, and we will confine our attention to this case hereafter.
	\item \textcolor{red}{(TODO)}: Introduce the Sobolev embedding theorem, for at least three reasons: (1) to facilitate later discussion of why smoothing splines are ill-posed when $2s < d$; (2) to make explicit why the statistical problem isn't ill-posed when $2s < d$ in the random design setting; (3) to make clear why the testing problem is hopeless when $4s < d$ without imposing additional conditions.
\end{itemize}

\paragraph{(2.b) Connecting Laplacian smoothing to smoothing splines.} 
To better motivate Laplacian smoothing as a tool well-suited for nonparametric regression, we review their connection to one of the most classical methods of nonparametric regression: smoothing splines.
\begin{itemize}
	\item Smoothing splines are particularly well-suited for nonparametric regression over Sobolev spaces. This is because smoothing splines are penalized least squares estimators where the penalty is itself the Sobolev semi-norm. Mathematically, the smoothing spline estimator $\wh{f}_{\SM}$ is defined as the solution to the following optimization problem,
	\begin{equation}
	\label{eqn:smoothing_spline}
	\wt{f} = \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \cdot \seminorm{f}_{H^{1}(\Xset)}^2\biggr\}.
	\end{equation}
	In~\eqref{eqn:smoothing_spline}, the minimum can either be taken over all $f \in H^1(\Xset)$ or over all $f \in C^1(\Xset)$; we shall refer to the former as smoothing splines, and the latter as \emph{Holder} smoothing splines.

	\textcolor{red}{(TODO)}: Check the classical spline references, and/or Ryan, for data on (1) what the canonical domain of minimization is for smoothing splines, and (2) whether the representer theorem depends holds for either domain of minimization.

	\item By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can tie the penalty used by the smoothing spline estimator $\wh{f}_{\SM}$ to that used by the Laplacian smoothing estimator $\wh{f}_{\LS}$. Suppose for the moment that $P$ is the uniform distribution over $\Xset$. Using the definition~\eqref{eqn:laplace_operator}, integrating by parts, and invoking the aforementioned convergence properties of $\Lap_{n,r}$, we have that for any $f \in C^{2}(\Xset)$ which satisfies appropriate boundary conditions,
	\begin{equation}
	\label{eqn:seminorm_convergence}
	\abs{f}_{H^1(\Xset)} = \int_{\Xset} \bigl(\Delta_P f\bigr)(x) \cdot f(x) \,dx = \lim_{n \to \infty} \frac{1}{n^{2}r^{(d + 2)}}f^T \Lap_{n,r} f.
	\end{equation}
	\textcolor{red}{(TODO)}: Formalize the boundary conditions---I believe either Neumann or Dirichlet boundary conditions should suffice?
	\item \eqref{eqn:seminorm_convergence} establishes asymptotic equality (up to a scaling factor) between the penalties in~\eqref{eqn:smoothing_spline} and~\eqref{eqn:laplacian_smoothing}, when the distribution $P$ is uniform. But it also sheds light on additional feature of Laplacian smoothing, relevant when $P$ is non-uniform: the penalty $f^T \Lap_{n,r} f$ automatically adapts to the design density $p$, in a way that the penalty $|f|_{H^1(\Xset)}$ does not. This automatic adaptivity to $P$ turns out to be a formal advantage of Laplacian smoothing, which we will comment on further in Section~\ref{sec:minimax_optimal_laplacian_smoothing}.
\end{itemize}

\paragraph{(2.c) Minimax-optimality of smoothing splines.} 
One reason all of this is intriguing is that, as mentioned previously, the smoothing spline estimator $\wh{f}_{\SM}$---and the resultant test statistic $T_{\SM} := \|\wt{f}_{\SM}\|_n^2$--- have strong statistical properties. To be precise, they achieve the well-known minimax rates over Sobolev spaces.
\begin{itemize}
	\item  The minimax estimation rate over Sobolev balls, with error measured in $\Leb^2(\Xset)$ norm, is 
	\begin{equation}
	\label{eqn:sobolev_space_estimation_minimax_rate}
	\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{d/(2 + d)}n^{-2/(2 + d)}~~\textrm{for all $d \geq 1$.}
	\end{equation}

	\textcolor{red}{(TODO)}: (1) You need a reference for this statement, and you already know one will be tricky to find because of ill-definedness of point evaluation. (2) This statement only holds under some condition on the density $p$.

	\item On the other hand, the minimax critical radius over Sobolev balls---that is, the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has non-trivial power over all $H_{\epsilon}^1(\Xset,M) := H^1(\Xset,M) \cap \{f: \norm{f}_{\Leb^2(\Xset)} \geq \epsilon\}$---is much smaller, 
	\begin{equation}
	\label{eqn:sobolev_space_testing_critical_radius}
	\epsilon_n^{\star}\Bigl(H^1(\Xset,M)\Bigr):= \inf\Biggl\{\epsilon: \inf_{\phi} \biggl[\Ebb_0[\phi] +  \sup_{f_0 \in H_{\epsilon}^1(\Xset,M)} \Ebb_{f_0}[1 - \phi]\biggr] \leq 2\alpha \Biggr\} \asymp n^{-2/(4 + d)}~~\textrm{for $1 \leq d < 4$,}
	\end{equation}

	\textcolor{red}{(TODO)}: (1) You need a reference for this statement (Ingster), and (2) you need to explain what happens when $d \geq 4$.

	\item The estimator $\wh{f}_{\SM}$ and the level-$\alpha$ test $\phi_{\SM} := \1\{T_{\SM} \geq {t}_{\alpha}\}$ are known to be minimax rate-optimal---meaning they achieve the rates in~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius}, respectively---when $d = 1$. When $d > 1$, as a consequence of the Sobolev Embedding Theorem\footnote{Or to be more precise, as a consequence of the \emph{tightness} of the Sobolev Embedding Theorem.} the variational problem~\eqref{eqn:smoothing_spline} is not well-posed: the criterion is minimized by interpolating the responses $\mathbf{Y}$, and the resulting estimator $\wt{f}$ will be inconsistent. By replacing the domain of minimization $H^1(\Xset)$ by $C^1(\Xset)$, we can force the solution to~\eqref{eqn:smoothing_spline} to be less wild, and the resulting Holder smoothing spline is not merely consistent but in fact minimax optimal when $d = 2$.\footnote{At least in the estimation problem. We could not find either a positive or negative result for the testing problem.} However when $d > 2$, the Holder space $C^1(\Xset)$ is itself too large, and even the Holder smoothing spline has risk provably larger than optimal. Previewing things a bit, we will see in Section~\ref{sec:minimax_optimal_laplacian_smoothing} that a similar phenomenon occurs for the Laplacian smoother $\wh{f}$; interestingly, the breakdown point will come when $d > 4$, rather than $d > 1$ or $d > 2$.

	\textcolor{red}{(TODO)}: Include references.

	\item Finally, it is well known that if $\Xset$ satisfies the manifold hypothesis, meaning informally that it has intrinsic dimension $m < d$ (see~\ref{asmp:domain_manifold} in Section~\ref{sec:graph_sobolev_classes} for a formal statement), both~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$. Encouragingly, the smoothing spline estimator $\wt{f}$ and test $\wt{\phi}$ achieve these faster rates, with one caveat: in order for smoothing splines to properly leverage the manifold hypothesis, the domain $\Xset$ must be known.
\end{itemize}

\paragraph{(2.d) Consequences?} Intuitively, the connections between Laplacian smoothing and smoothing splines give us reason to hope that the strong statistical properties of the latter might be shared by the former. Unfortunately, it is not so easy to mathematically tie $\wh{f}$ to $\wt{f}$: for example, note that the convergence in~\eqref{eqn:seminorm_convergence} assumes $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:smoothing_spline} is either $H^1(\Xset)$ or $C^1(\Xset)$. Instead, we will take a different approach and analyze $\wh{f}$ by leveraging properties of the graph Sobolev class $H^1(G_{n,r})$, which we now introduce.

\section{Graph Sobolev classes}
\label{sec:graph_sobolev_classes}

\textbf{(3)}  For $M > 0$, define the (first-order) graph Sobolev ball to be
\begin{equation}
\label{eqn:graph_sobolev_ball}
H^1(G_{n,r},M) = \Bigl\{f \in \Reals^n: f^T \Lap_{n,r} f \leq M^2\Bigr\}
\end{equation}
where we recall that the \emph{graph Sobolev seminorm} $f^T \Lap_{{n,r}} f$ is precisely the penalty term in~\eqref{eqn:laplacian_smoothing}.

To understand why graph Sobolev balls play a critical role in the analysis of Laplacian smoothing, let us return for a moment to the continuous-time setting. Upper bounds on the error of smoothing splines rest on two crucial facts, both of which involve the continuum Laplace operator $\Delta_P$. The first is the \textit{a priori} assumption that the Sobolev semi-norm $|f|_{H^1(\Xset)} = \langle f, \Delta_P f \rangle_{\Leb^2(P)} \leq M^2$.\footnote{The equality holding when $P$ is uniform and  only for functions $f$ which satisfy boundary conditions.} The second is an upper bound on the metric entropy of $H^1(\Xset)$, characterized by the eigenvalue decay $\lambda_k(\Delta_P) \asymp k^{2/d}$ (i.e. Weyl's Law). In this section we will show that both these statements have analogues in the graph setting:
\begin{itemize}
	\item for functions $f$ belonging to $H^1(\Xset,M)$, the evaluations $f \in (f(X_1),\ldots,f(X_n)) \in \Reals^n$ are upper bounded,
	\begin{equation}
	\label{eqn:graph_sobolev_seminorm_vague}
	f^T \Lap_{n,r} f \leq C M^2(n^{2}r^{(d + 2)}),
	\end{equation}
	\item for a range $1 \leq k \leq k_{\max}(n)$, the eigenvalues $\lambda_k(G_{n,r})$ satisfy
	\begin{equation}
	\label{eqn:neighborhood_graph_eigenvalue_vague}
	ck^{2/d} \leq \frac{1}{nr^{d + 2}} \lambda_k(G_{n,r})  \leq Ck^{2/d}.
	\end{equation}
\end{itemize}
Both statements will hold with high probability, and when we assume $\Xset$ is a manfiold of dimension $m$, they will hold with $d$ replaced by $m$.  These results elucidate why Laplacian smoothing ``works'' as well as its continuous-time counterpart ---in the sense of being statistically rate-optimal over (continuum) Sobolev classes.

\paragraph{(3.a) Graph Sobolev semi-norm.}
We begin by upper bounding $f_0^T \Lap_{n,r} f_0$ under the assumption $f_0 \in H^1(\Xset,M)$. We will assume the kernel $K$ satisfies condition~\ref{asmp:kernel}.
\begin{enumerate}[label=(K\arabic*)]
	\item
	\label{asmp:kernel}
	$K:[0,\infty) \to [0,\infty)$ is supported on $[0,1]$, and its restriction to $[0,1]$ is Lipschitz. Additionally, it is normalized so that
	\begin{equation*}
	\int_{\Reals^d} K\bigl(\norm{z}\bigr) \,dz = 1.
	\end{equation*}
\end{enumerate}
This is a relatively mild assumption: the choice of kernel is under the user's control, and moreover~\ref{asmp:kernel} covers many (though certainly not all) common kernel choices. In Lemma~\ref{lem:graph_sobolev_seminorm}, we prove that if the density $p$ is merely upper bounded then~\eqref{eqn:graph_sobolev_seminorm_vague} holds with respect to the graph Sobolev semi-norm.
\begin{lemma}
	\label{lem:graph_sobolev_seminorm}
	Assume~\ref{asmp:kernel} and additionally that $p(x) \leq p_{\max}$ for all $x \in \Xset$. Then, for any $f \in H^1(\Xset)$, there exists a constant $C_1 > 0$ which depends only on $\Xset$, $d$, $p_{\max}$ and $K$ such that
	\begin{equation}
	\label{eqn:graph_sobolev_seminorm}
	f^T \Lap_{n,r} f \leq \frac{C_1}{\delta} n^2 r^{d + 2} |f|_{H^1(X)}^2
	\end{equation}
	with probability at least $1 - \delta$.
\end{lemma}
A few remarks:
\begin{itemize}
	\item 
	The inequality~\eqref{eqn:graph_sobolev_seminorm} resembles~\eqref{eqn:seminorm_convergence}. However, as we pointed out previously to derive~\eqref{eqn:seminorm_convergence} one needs $f \in C^{2}(\Xset)$ as opposed to $f \in H^1(\Xset)$, and imposing this extra degree of regularity is unsuitable for our purposes. We use a more careful analysis to prove Lemma~\eqref{lem:graph_sobolev_seminorm}, but note well that~\eqref{eqn:graph_sobolev_seminorm} is only an upper bound rather than a statement about consistency. This is the price we pay for insisting that $f$ have only one derivative; fortunately an upper bound will be sufficient for our purposes in Section~\ref{sec:minimax_optimal_laplacian_smoothing}.
	\item We note also that~\eqref{eqn:graph_sobolev_seminorm}---which holds with probability $1 - \delta$---depends proportionally on $1/\delta$, as opposed to $\log(1/\delta)$. The reason: the Sobolev assumption $f \in H^1(\Xset)$ only grants us control over the second moment of $f$ and its first derivative, and so we are afforded only the relatively weak concentration implied by Markov's inequality. If we assume $f$ and its first derivative are more uniformly bounded--for instance if $f \in C^1(\Xset)$---we can use Hoeffding's inequality to get sharper upper bounds.
\end{itemize}

\textbf{(3.b) Graph Laplacian eigenvalues.} The graph eigenvalues $\lambda_k(\Lap_{n,r}) \asymp n r^{d + 2}k^{2/d}$---when $\Xset$ is full-dimensional---or $\lambda_k(\Delta_P) \asymp nr^{m + 2}k^{2/m}$---when $\Xset$ has intrinsic dimension $1 < m < d$.

\section{Minimax optimality of Laplacian smoothing}
\label{sec:minimax_optimal_laplacian_smoothing}

\textbf{(4)} We now leverage the results of Section~\ref{sec:graph_sobolev_classes} to formalize the main conclusions of this work: that Laplacian smoothing methods are minimax rate-optimal.

\textbf{(4.a) Estimation error of Laplacian smoothing.} Using \textbf{(3.a)} and \textbf{(3.b)}, we establish upper bounds on the estimation error of Laplacian smoothing.

\textbf{(4.b) Testing error of Laplacian smoothing.} Using \textbf{(3.a)} and \textbf{(3.b)}, we establish upper bounds on the testing error of Laplacian smoothing.

\quad \textbf{(4.b.i) Low-smoothness regime.} Unlike in the estimation setting, the nonparametric goodness-of-fit testing problem itself exhibits an elbow when $d = 4$ and $f_0 \in \Leb^4(\Xset)$. The naive statistic $\frac{1}{n}\|Y\|_2^2$ is now optimal. By tuning Laplacian smoothing to the interpolation limit---i.e. setting $\rho = 0$---we recover this naive statistic, and therefore Laplacian smoothing is optimal in this sense (although really, no smoothing is being done.)

\textbf{(4.c) Comparing with smoothing splines.} While we motivated graph Laplacians by suggesting they might replicate the statistical properties of smoothing splines, we have in fact managed to show something stronger. In particular:

\quad \textbf{(4.c.i)} When $d \geq 2$, there is no Sobolev embedding of $H^1(\Xset) \subseteq C^s(\Xset)$ for any $s > 0$. Therefore the smoothing spline estimator is ill-posed.

\quad \textbf{(4.c.ii)} Additionally, the metric entropy of $C^1(\Xset)$ is too large to obtain optimal rates for \textcolor{red}{Holder-smoothing splines} when $d \geq 2$. Thus, we see a gap emerge between the optimality properties of the discrete-time (Laplacian smoothing) and continuous-time (smoothing splines) estimators, in favor of the latter.

\quad \textbf{(4.c.iii)} On the other hand, when $d \geq 4$, we no longer get the known optimal rates in the estimation problem. When $d = 4$, our rates are suboptimal by a factor of $\log n$. When $d > 5$ they are very suboptimal. We believe this is related to classical results regarding the entropy of low-smoothness Holder classes, i.e. the results used to justify \textbf{(4.c.ii)}. The interesting point is that the parameter spaces $H^1(\Xset)$, $C^1(\Xset)$ and $H^1(G_{n,r})$ become ``too large'' for different values of $d$.

\quad \textbf{(4.c.iv)} \textcolor{red}{(TODO)}: Translate these results into error in $\Leb^2(P)$ error, if/when possible.

\textbf{(4.d) Manifold adaptivity.} In Section~\ref{sec:graph_sobolev_classes}, we established that when $\Xset$ is an $m$-dimensional submanifold of $\Reals^d$, Laplacian smoothing approaches automatically ``feel'' the intrinsic dimension $m$ of $\Xset$. We now show that this translates into improved upper bounds on estimating and testing error.

\quad \textbf{(4.d.i)} This confirms the statement in \textbf{(4)} that Laplacian smoothing automatically adapts to the geometry of $\Xset$ and $p$.

\quad \textbf{(4.d.ii)} By contrast, smoothing splines must be specially tailored to work in this setting.

\textbf{(4.e) Higher-order smoothness classes.} Laplacian smoothing can also be adapted to take advantage of additional assumed regularity on $f_0$, i.e. $f_0 \in H^s(\mc{\Xset})$ for $s > 1$. In the very special case of $s = 2$ and $d = 2$, we can show that this adapted estimator achieves the sharper minimax rates over these classes, but the general story--for all combinations of $s$ and $d$---remains beyond our reach.

\section{Simulations}

\textbf{(5)} Empirically, we demonstrate that the risk for Laplacian smoothing is comparable to that of smoothing splines---in the full dimensional case with uniformly random design. When the design is sampled from a density concentrated near a manifold, Laplacian smoothing outperforms (out of the box) smoothing splines.

\section{Discussion}

\textcolor{red}{(TODO)}

\clearpage

\appendix

\clearpage

Many of our results follow the same general two-part proof strategy. First, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to functionals of the graph $G$. We then analyze the behavior of these functionals with respect to the particular neighborhood graph $G_{n,r}$ and give high probability (upper or lower) bounds on these functionals. It is in this second step where we invoke our various assumptions on the distribution $P$ and regression function $f_0$.

\section{Graph-dependent error bounds}
\label{sec:fixed_graph_error_bounds}

\section{Neighborhood graph Sobolev semi-norm}
\label{sec:graph_sobolev_seminorm}

\section{Bounds on neighborhood graph eigenvalues}
\label{sec:graph_eigenvalues}

\section{Bounds on the empirical norm}
\label{sec:empirical_norm}

\section{Proof of theorems}

\bibliographystyle{plainnat}
\bibliography{../../../graph_testing_bibliography}

\end{document}

### 2020-09-30 notes/weekly_notes/9.23.20.NotesForWeek.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{tikz}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\iid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}
\newcommand{\Lap}{L}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}
\newcommand{\Id}{I}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Vset}{\mathcal{V}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 9/16/2020 - 9/23/2020}
\author{Alden Green}
\date{\today}
\maketitle

In these notes, I will carefully examine the results of \textcolor{red}{(GarciaTrillos18, Calder19)} on the convergence of neighborhood-graph eigenvalues towards eigenvalues of a continuum Laplace-Beltrami operator. The idea is to: (a) prove results in the ``flat Euclidean'' setting (where the manifold is full-dimensional and has boundary), and (b) to ensure there are no hidden dependencies in the stated constants that I am missing.

\section{Main results}
Let $X_1,\ldots,X_n$ be sampled iid from a distribution $P$, supported on a compact set $\Xset \subset \Reals^d$.

\begin{enumerate}[(P1)]
	\item 
	\label{asmp:bounded_lipschitz_density} We assume the distribution $P$ admits density $p$ that is Lipschitz continuous with Lipschitz parameter $L_p$ and satisfies
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty,~~\textrm{for all $x \in \Xset$.}
	\end{equation*}
\end{enumerate}

For a given radius $r > 0$, and kernel function $K: [0,\infty) \to [0,\infty)$, define the neighborhood graph $G_{n,r} = ([n],\mathbf{W})$ to be the graph with weighted edges $\mathbf{W}_{ij} = K(\|X_i - X_j\|/r)$. The neighborhood graph Laplacian is $\Lap_{n,r} = \mathbf{D} - \mathbf{W}$, where $\mathbf{D}$ is the degree matrix, and $0 = \lambda_1(G_{n,r}) \leq \cdots \leq \lambda_n(G_{n,r})$ are the eigenvalues of $\Lap_{n,r}$. Our goal is to relate these eigenvalues to the eigenvalues $\lambda_k(\Delta_P)$ of the Laplace-Beltrami operator
\begin{equation*}
\Delta_{P}f  := -\frac{1}{2p} \dive(p^2 \nabla f)
\end{equation*}
We would like to show the following result.
\begin{theorem}
	\label{thm:neighborhood_eigenvalue}
	There constants $A_1$ and $a_1$ which do not depend on $n$ or $r$ such that the following statement holds with probability at least $1 - A_1n\exp(-a_1n\theta^2\wt{\delta}^{m})$: for any $\ell \in \mathbb{N}$ such that
	\begin{equation}
	\label{eqn:neighborhood_eigenvalue_1}
	1 - A_1\Biggl(r \sqrt{\lambda_{\ell}(\Delta_P)} + (\theta + \wt{\delta})\Biggr)  \geq \frac{1}{2}
	\end{equation}
	we have that
	\begin{equation}
	\label{eqn:eigenvalue_bound}
	\frac{1}{A_1} \lambda_k(G_{n,r}) \leq nr^{d+2} \lambda_k(\Delta_P) \leq A_1 \lambda_k(G_{n,r}),~~\textrm{for all $1 \leq k \leq \ell$}
	\end{equation}
\end{theorem}
Some results along these lines (see \textcolor{red}{(Burago13, GarciaTrillos18, Calder19)}) have been shown when $\Xset$ is an $m$-dimensional submanifold of $\Reals^d$---in this case, replace $r^{d + 2}$ by $r^{m+2}$ in~\eqref{eqn:eigenvalue_bound}. Indeed in these results the constant $A = (1 + e)$ for an error term $e$ which goes to $0$ as $n \to \infty, r(n) \to 0$. However, (a) the manifold is assumed to be boundaryless, and (b) the right hand side in (their equivalent of)~\eqref{eqn:neighborhood_eigenvalue_1} depends on $\ell$ which is a bit strange and unsuitable for our purposes. We would like to ``fix'' both (a) and (b).

\section{Ancillary results.}
Following the lead of~\textcolor{red}{(Burago13, GarciaTrillos18, Calder19)}, we relate $\lambda_k(\Delta_P)$ and $\lambda_k(G_{n,r})$ by means of the Sobolev semi-norms (Dirichlet energies)
\begin{equation*}
b_r(f) := \frac{1}{n^2 r^{d+ 2}}f^T \Lap_{n,r} f~~\textrm{for any $f \in \Leb^2(P_n)$}
\end{equation*}
and
\begin{equation*}
D_2(f) :=
\begin{cases*}
\int_{\Xset} \|\nabla f(x)\|^2 p^2(x) \,dx~~ &\textrm{if $f \in H^1(\Xset)$} \\
\infty~~ & \textrm{otherwise,}
\end{cases*}
\end{equation*}
Suppose one could upper bound $b_r(u)$ by $D_2\bigl(G(u)\bigr)$ for an appropriate choice of mapping $G: \Leb^2(P_n) \to \Leb^2(P)$, and vice versa upper bound $D_2(f)$ by $b_r(V(f))$ for some $V: \Leb^2(P) \to \Leb^2(P_n)$. Suppose also that $G$ and $V$ are near-isometries, meaning $\|G(u)\|_{\Leb^2(P)} \approx \|u\|_{\Leb^2(P_n)}$ and $\|V(f)\|_{\Leb^2(P_n)} \approx \|f\|_{\Leb^2(P)}$. Then using the variational characterization of eigenvalues $\lambda_k(\Delta_P)$ and $\lambda_k(G_{n,r})$---i.e. the Courant-Fischer Theorem---one can obtain estimates on the error of $|\lambda_k(\Delta_P) - \lambda_k(G_{n,r})|$.

We now move to define maps $V$ and $G$ which provably satisfy such bounds. Let $\wt{P}_n$ be a probability measure with density $\wt{p}_n$, that satisfy the conclusions of Theorem~\ref{thm:calder19_2}. Denote by $\wt{T}_n$ an \emph{optimal transport map} (in sup-norm) between $\wt{P}_n$ and $P_n$, and let $U_1,\ldots,U_n$ be the preimages of $X_1,\ldots,X_n$ under $\wt{T}_n$. Define the discretization map  $\wt{\mathcal{P}}: \Leb^2(P) \to \Leb^2(P_n)$ by
\begin{equation*}
\bigl(\wt{\mathcal{P}}f\bigr)(X_i) := n \cdot \int_{U_i} f(x) \wt{p}_n(x) \,dx
\end{equation*}
Define the interpolation map $\wt{\mc{I}}: \Leb^2(P_n) \to \Leb^2(\Xset)$ as $\wt{\mc{I}}u := \Lambda_{r - 2\wt{\delta}}(\wt{\mc{P}}^{\star}u)$. Here, $\wt{\mc{P}}^{\star} = u \circ \wt{T}$ is the adjoint of $\wt{\mc{P}}$, i.e.
\begin{equation*}
\bigl(\wt{\mc{P}}^{\star}u\bigr)(x) = \sum_{j = 1}^{n} u(x_i) \1\{x \in U_i\} 
\end{equation*} 
and $\Lambda_r$ is a smoothening of this extension, formally
\begin{equation*}
\Lambda_r(f) := \frac{1}{r^d\tau(x)}\int_{\Xset} \eta_r(x',x) f(x') \,dx',~~ \eta_r(x',x) := \psi\biggl(\frac{\|x' - x\|}{r}\biggr)
\end{equation*}
where $\psi(t) := (1/\sigma_K)\int_{t}^{\infty} s K(s) \,ds$ and $\tau(x) := (1/r^d)\int_{\Xset} \eta_r(x',x) \,dx'$ is a normalizing constant.

Assume the following.
\begin{enumerate}[label=(K\arabic*)]
	\item
	\label{asmp:kernel}
	$K:[0,\infty) \to [0,\infty)$ is a non-increasing function supported on $[0,1]$, and its restriction to $[0,1]$ is Lipschitz with Lipschitz constant $L_K$. Additionally, it is normalized so that
	\begin{equation*}
	\int_{\Reals^d} K\bigl(\norm{z}\bigr) \,dz = 1,
	\end{equation*}
	and $K(1/2) > 0$. Let $\sigma_{K} := \frac{1}{d}\int_{\Reals^d} \|z\|^2 K(\|z\|) \,dy$.
\end{enumerate}
\begin{enumerate}[label=(OT\arabic*)]
	\item 
	\label{asmp:calder19_1}
	There exists $n^{-1/m} < \wt{\delta} < \frac{1}{4}r$.
	\item 
	\label{asmp:calder19_2} There exists $\theta$ such that $C(\theta + \wt{\delta}) \leq p_{\min}/2$.
	\item
	\label{asmp:r_small_1} For all $x,x' \in \Xset$ such that $\|x - x'\| \leq r$, we have that for $W = B(x,r) \cap B(x',r) \cap \Xset$,
	\begin{equation*}
	\vol(W) \geq a_1(\Xset,d) \cdot \nu_d \biggl(\frac{r}{2}\biggr)^d
	\end{equation*}
	for some constant $A_1(\Xset,d)$. 
	\item
	\label{asmp:r_small_2} For all $x,x' \in \Xset$, we have that $z_{xx'} = \frac{1}{2}x + \frac{1}{2}x' \in \Xset$.
	\item 
	\label{asmp:r_small_3} For any $x \in \Xset$, letting $\mc{B}_x:= (\Xset - x)/r \cap B(0,1)$, we have that $\int_{\mc{B}_x} \psi(\|z\|) \,dz \geq a_1(\Xset,d) \cdot \int_{B(0,1)} \psi(\|z\|) \,dz$.
\end{enumerate}
Very roughly speaking, one can view $\theta + \wt{\delta}$ as an upper bound on the optimal transport distance between $P_n$ and $\wt{P}_n$ and the $\Leb_{\infty}$ distance between $\wt{p}_n$ and $p$.

\textcolor{red}{(TODO)}: Eliminate these assumptions, by showing that each of them is a conseequence of the regularity assumption that $\Xset$ is a Lipschitz domain with positive reach.

\begin{proposition}[\textbf{c.f. Proposition 4.1 of \textcolor{red}{Calder19}}]
	\label{prop:dirichlet_energies}
	With probability at least $1 - A_1n\exp(-a_1n\theta^2\wt{\delta}^{m})$, we have the following.
	\begin{enumerate}[(1)]
		\item For any $u \in \Leb^2(P_n)$,
		\begin{equation*}
		\sigma_{K} D_2(\wt{\mc{I}}u) \leq \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \biggl(1 + A_3(d,K,p)\frac{\wt{\delta}}{r}\biggr) b_r(u)
		\end{equation*}
		\item For any $f \in \Leb^2(P)$,
		\begin{equation*}
		b_r(\wt{\mc{P}}f) \leq \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + L_p(2r + 4\wt{\delta})\Bigr) \cdot \Bigl(1 + A_1(K)\frac{\wt{\delta}}{r}\Bigr) \sigma_{K} D_2(f).
		\end{equation*}
	\end{enumerate}
\end{proposition}

\begin{proposition}[\textbf{c.f. Proposition 4.2 of \textcolor{red}{Calder19}}]
	\label{prop:isometry}
	With probability at least $1 - A_1n\exp(-a_1n\theta^2\wt{\delta}^{m})$, we have the following.
	\begin{enumerate}[(1)]
		\item For every $f \in \Leb^2(\Xset)$,
		\begin{equation*}
		\Bigl|\|f\|_{\Leb^2(P)}^2 - \|\wt{P}f\|_{\Leb^2(P_n)}^2\Bigr| \leq A_2(\Xset,d,K,p_{\min}) \cdot \wt{\delta} \|f\|_{\Leb^2(P)} \sqrt{D_2(f)} + \frac{A_1\bigl(\theta + \wt{\delta}\bigr)}{p_{\min}} \|f\|_{\Leb^2(P)}^2
		\end{equation*}
		\item For every $u \in \Leb^2(P_n)$,
		\begin{equation*}
		\Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(P)}^2 - \|u\|_{\Leb^2(P_n)}^2\Bigr| \leq C r \|u\|_{\Leb^2(P_n)} \sqrt{b_r(u)} + C\bigl(\theta + \wt{\delta}\bigr) \|u\|_{\Leb^2(P_n)}^2
		\end{equation*}
	\end{enumerate}
\end{proposition}

To prove these inequalities, it will help to introduce the intermediate energies
\begin{equation*}
\wt{E}_r(f,\eta,V) := \frac{1}{r^{d + 2}}\int_{V} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{r}\biggr) \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx
\end{equation*}
and
\begin{equation*}
{E}_r(f,\eta,V) := \frac{1}{r^{d + 2}}\int_{V} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx
\end{equation*}
where $\eta: [0,\infty) \to [0,\infty)$ is an arbitrary kernel, and $V \subseteq \Xset$ is a measurable set. We will abbreviate $\wt{E}_r(f,\eta,\Xset)$ as $\wt{E}_r(f,\eta)$ and $\wt{E}_r(f,K) = \wt{E}_r(f)$ (and likewise with $E_r$.)

Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized} relates the graph Sobolev semi-norm $b_r(\wt{\mc{P}}f)$ to the non-local energy $\wt{E}_r(f)$. It follows exactly from the derivations in the proof of Lemma 13 in \cite{trillos2019}; indeed, in our flat Euclidean setting, the proof is simplified, and we include this simpler proof for completeness purposes only.
\begin{lemma}[\textbf{c.f. Lemma 13 of \cite{trillos2019}, Lemma 4.3 of \cite{burago2014}}]
	\label{lem:first_order_graph_sobolev_seminorm_discretized}
	Suppose~\ref{asmp:kernel} is satisfied, and that the conclusions of Theorem~\ref{thm:calder19_2} are met. Then
	\begin{equation*}
	b_r(\wt{\mc{P}}f) \leq \wt{E}_{r + 2\wt{\delta}}(f) + 2\frac{L_K\wt{\delta}}{K(1/2)r} \wt{E}_{2(r + 2\wt{\delta})}(f)
	\end{equation*}
	for any $f \in \Leb^2(\Xset)$.
\end{lemma}

Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected} relates the non-local energy $E_r(f)$ to $D_2(f)$. It follows in large part from the proofs of Lemma 13 of \cite{trillos2019} and Lemma 6 of \cite{burago2014}, with minor modifications to account for the boundary of $\Xset$.
\begin{lemma}[c.f. Lemma 13 of \cite{trillos2019}, Lemma 6 of \cite{burago2014}]
	\label{lem:first_order_graph_sobolev_seminorm_expected}
	For any $f \in \Leb^2(\Xset)$,
	\begin{equation*}
	E_r(f) \leq (1 + L_pr)^2 \cdot \sigma_K D_2(f)
	\end{equation*}
\end{lemma}

In Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, we establish the reverse of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected}. 
\begin{lemma}[c.f. Lemma 9 of \cite{trillos2019}, Lemma 5.5 of \cite{burago2014}]
	\label{lem:first_order_graph_sobolev_seminorm_expected_lb}
	For any $f \in \Leb^2(\Xset)$,
	\begin{equation*}
	\sigma_KD_2(\Lambda_rf) \leq \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K}E_r(f)
	\end{equation*}
\end{lemma}

In Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}, we establish the reverse of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}. 
\begin{lemma}[c.f. Lemma 14 of \cite{trillos2019}]
	\label{lem:first_order_graph_sobolev_seminorm_discretized_lb}
	Suppose $r$ satisfies~\ref{asmp:r_small_2}. For any $u \in \Leb^2(P_n)$, 
	\begin{equation*}
	\wt{E}_{r - 2\wt{\delta}}\bigl(\wt{\mc{P}}^{\star}u\bigr) \leq \biggl(1 + A_3(d,K,P)\frac{\wt{\delta}}{r}\biggr) b_{r}(u)
	\end{equation*}
\end{lemma}

To prove Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, we require upper and lower bounds on $\tau(x)$, as well as an upper bound on the gradient of $\tau$. The lower bound here---$\tau(x) \geq 1/2$---is quite a bit a looser than in the case where $\Xset$ has no boundary---$\tau(x) \geq (1 + Cr^2)^{-1}$. The same is the case regarding the upper bound of the size of the gradient $\|\nabla \tau(x)\|$. However, the bounds as stated here will be sufficient for our purposes.
\begin{lemma}
	\label{lem:tau_bound}
	Suppose $r$ satisfies~\ref{asmp:r_small_3}. For all $x \in \Xset$,
	\begin{equation*}
	\frac{1}{2} \leq \tau(x) \leq 1,
	\end{equation*}
	and additionally $\|\nabla \tau(x)\| \leq \frac{1}{\sigma_K r}$.
\end{lemma}

To prove part (2) of Proposition~\ref{prop:isometry}, we require Lemma~\ref{lem:smoothening_error}, which gives an estimate on the error $\Lambda_h f - f$ in $\Leb^2(P)$ norm.
\begin{lemma}[c.f Lemma 8 of \cite{trillos2019}, Lemma 5.4 of \cite{burago2014}]
	\label{lem:smoothening_error}
	For any $h > 0$, 
	\begin{equation}
	\label{eqn:smoothening_error_norm}
	\bigl\|\Lambda_hf\bigr\|_{\Leb^2(P)}^2 \leq \frac{2p_{\max}}{p_{\min}} \bigl\|f\bigr\|_{\Leb^2(P)}^2
	\end{equation}
	and
	\begin{equation}
	\label{eqn:smoothening_error_energy}
	\bigl\|\Lambda_hf - f\bigr\|_{\Leb^2(P)}^2 \leq \frac{2}{\sigma_Kp_{\min}} r E_r(f)
	\end{equation}
	for all $f \in \Leb^2(\Xset)$.
\end{lemma}

\section{Other's results}

For a transport map $T: \Xset \to \mathbf{X}$ and a measure $\mu$ supported on $\Xset$, let $T_{\sharp}\mu$ denote the push-forward of $\mu$ by $T$, i.e the measure for which
\begin{equation*}
\bigl(T_{\sharp}\mu\bigr)(\mathbf{Z}) = \mu\bigl(T^{-1}(\mathbf{Z}))
\end{equation*}
for any $\mathbf{Z} \subseteq \mathbf{X}$.

\begin{theorem}[Proposition~2.11 of \textcolor{red}{(Calder2019)}]
	\label{thm:calder19_2}
	Suppose the assumptions~\ref{asmp:calder19_1} and~\ref{asmp:calder19_2} are satisfied. Then with probability at least $1 - A_1n\exp(-a_1n\theta^2\wt{\delta}^d)$, there exists a probability measure $\wt{P}_n$ with density function $\wt{p}_n: \Xset \to \Reals$ such that
	\begin{equation*}
	\min_{T_{\sharp}\wt{P}_n = P_n} d_{\mc{M}}(x,T(x)) \leq \wt{\delta}
	\end{equation*}
	and
	\begin{equation*}
	\bigl\|\wt{p}_n - p\bigr\|_{\infty} \leq A_1\bigl(\theta + \wt{\delta}\bigr)
	\end{equation*}
\end{theorem}
\textcolor{red}{(TODO)}: Replace this Theorem with it's ``flat'' predecessor.

By Theorem~\ref{thm:calder19_2}, the two energies can be related,
\begin{equation}
\label{eqn:calder19_1}
\bigl(1 - A_1(\theta + \wt{\delta})\bigr) E_r(f) \leq \wt{E}_r(f) \leq \bigl(1 + A_1(\theta + \wt{\delta})\bigr) E_r(f),
\end{equation}
and so can the norms
\begin{equation}
\label{eqn:calder19_2}
\bigl(1 - A_1(\theta + \wt{\delta})\bigr) \|f\|_{\Leb^2(P)}^2 \leq \|f\|_{\Leb^2(\wt{P}_n)}^2 \leq \bigl(1 + A_1(\theta + \wt{\delta})\bigr) \|f\|_{\Leb^2(P)}^2.
\end{equation}

\section{Additional proofs}

\paragraph{Proof of Theorem~\ref{thm:neighborhood_eigenvalue}.}
\mbox{} \\
\mbox{} \\
We start with the upper bound, proceeding exactly as in \citep{burago2014} (Proposition 4.4). Let $W$ be the span of $f_1,\ldots,f_{\ell}$, the first ${\ell}$ eigenfunctions of $-\Delta_P$, so that for every $f \in W$ we have $D_2(f) \leq \lambda_k(\Delta_P) \|f\|_{\Leb^2(P)}$. As a result, by Part (1) of Proposition~\ref{prop:isometry} we have that for any $f \in W$
\begin{equation*}
\bigl\|\wt{\mc{P}}f\bigr\|_{\Leb^2(P_n)}^2 \geq \biggl(1 - A_2(\Xset,d,K,p_{\min}) \wt{\delta} \sqrt{\lambda_{\ell}(\Delta_P)} - \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr)\|f\|_{\Leb^2(P)}^2  \geq \frac{1}{2} \|f\|_{\Leb^2(P)}^2,
\end{equation*}
where the second inequality follows by assumption~\eqref{eqn:neighborhood_eigenvalue_1}.

Therefore $\wt{\mc{P}}$ is injective over $W$, and $\wt{\mc{P}}W$ has dimension $\ell$. This means we can invoke the Courant-Fischer Theorem, along with Proposition \ref{prop:dirichlet_energies}, and conclude that
\begin{align*}
\lambda_k(G_{n,r}) & \leq \max_{\substack{u \in \wt{\mc{P}}W \\ u \neq 0} } \frac{b_r(u)}{\|u\|_{\Leb^2(P_n)}^2} \\
& = \max_{\substack{f \in W \\ f \neq 0} } \frac{b_r(\wt{\mc{P}}f)}{\bigl\|\wt{\mc{P}}f\bigr\|_{\Leb^2(P_n)}^2} \\
& \leq \frac{\Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + L_p(2r + 4\wt{\delta})\Bigr) \cdot \Bigl(1 + A_1(K)\frac{\wt{\delta}}{r}\Bigr)}{1 - A_2(\Xset,d,K,p_{\min}) \wt{\delta} \sqrt{\lambda_{\ell}(\Delta_P)} - 2\frac{A_1(\theta + \wt{\delta})}{p_{\min}}} \sigma_K \lambda_{\ell}(\Delta_P),
\end{align*}
establishing the upper bound in~\eqref{eqn:eigenvalue_bound} upon proper choice of $A$.

\textcolor{red}{(TODO)}: Ask Siva how to `finish this proof off'.

\paragraph{Proof of Proposition~\ref{prop:dirichlet_energies}.}
\mbox{} \\
\mbox{} \\
Part (1) follows from
\begin{align*}
\sigma_K D_2(\Lambda_{r - 2\wt{\delta}} \wt{\mc{P}}^{\star}u) & \overset{(i)}{\leq} \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} E_{r - 2\wt{\delta}}(\wt{\mc{P}}^{\star}u) \\
& \overset{(ii)}{\leq} \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \wt{E}_{r - 2\wt{\delta}}(\wt{\mc{P}}^{\star}u) \\
& \overset{(iii)}{\leq} \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \biggl(1 + A_3(d,K,p)\frac{\wt{\delta}}{r}\biggr) b_r(u)
\end{align*}
where $(i)$ follows from Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, $(ii)$ follows from~\eqref{eqn:calder19_1}, and $(iii)$ follows from~Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}.

Part (2) follows from
\begin{align*}
b_r(\wt{\mc{P}}f) & \overset{(iv)}{\leq} \wt{E}_{r + 2\wt{\delta}}(f) + A_1(K)\frac{\wt{\delta}}{r}\wt{E}_{2(r + 2\wt{\delta})}(f)\\
& \overset{(v)}{\leq} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl({E}_{r + 2\wt{\delta}}(f) + A_1(K)\cdot\frac{\wt{\delta}}{r}{E}_{2(r + 2\wt{\delta})}(f)\Bigr) \\
& \overset{(vi)}{\leq} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + L_p(2r + 4\wt{\delta})\Bigr) \cdot \Bigl(1 + A_1(K)\frac{\wt{\delta}}{r}\Bigr) \cdot \sigma_{\eta} D_2(f)
\end{align*}
where $(iv)$ follows from Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}, $(v)$ follows from~\eqref{eqn:calder19_1}, and $(vi)$ follows from Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected}.

\paragraph{Proof of Proposition~\ref{prop:isometry}.}
\mbox{}\\
\mbox{}\\
\textit{Proof of (1).}
\mbox{}\\
\mbox{}\\
We begin by upper bounding $\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}$. By the Cauchy-Schwarz inequality and the bound on $\|\wt{p}_n - p\|_{\infty}$ in~\eqref{eqn:calder19_1},
\begin{align*}
\Bigl|\wt{P}f(X_i)\Bigr|^2 & = n^2 \Bigl|\int_{U_i} f(x) \wt{p}_n(x) \,dx\Bigr|^2 \\
& \leq n \int_{U_i} \bigl|f(x)\bigr|^2 \wt{p}_n(x) \,dx \\
& \leq n \cdot \biggl[\int_{U_i} \bigl|f(x)\bigr|^2 p(x) \,dx + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \int_{U_i} \bigl|f(x)\bigr|^2 p(x) \,dx\biggr]
\end{align*}
and summing over $i = 1,\ldots,n$, we obtain
\begin{equation}
\label{pf:prop_isometry_1}
\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}^2 \leq \biggl(1 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \bigl\|f\bigr\|_{\Leb^2(P)}^2.
\end{equation}
Now, noticing that $\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)} = \bigl\|\wt{P}^{\star}\wt{P}f\bigr\|_{\Leb^2(\wt{P}_n)}$, we can use the upper bound~\eqref{pf:prop_isometry_1} to show that
\begin{align}
\Bigl|\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}^2 - \bigl\|f\bigr\|_{\Leb^2(P)}^2\Bigr| & \leq \Bigl|\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}^2 - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2\Bigr| + \Bigl|\bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2 - \bigl\|f\bigr\|_{\Leb^2(P)}^2\Bigr| \nonumber \\
& \overset{(i)}{\leq} \Bigl|\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}^2 - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2\Bigr|  + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{\Leb^2(P)}^2 \\
& \overset{(ii)}{\leq} \biggl(2 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \Bigl|\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)} - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}\Bigr| \cdot \bigl\|f\bigr\|_{\Leb^2(P)} + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{\Leb^2(P)}^2 \nonumber \\
& \leq \biggl(2 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)} \cdot \bigl\|f\bigr\|_{\Leb^2(P)} + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{\Leb^2(P)}^2 \label{pf:prop_isometry_2}
\end{align}
where $(i)$ follows from~\eqref{eqn:calder19_2} and $(ii)$ follows from~\eqref{pf:prop_isometry_1}.

It remains to upper bound $\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)}^2$. Noting that $\wt{P}^{\star}\wt{P}f$ is piecewise constant over the cells $U_i$, we have
\begin{equation*}
\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)}^2 = \sum_{i = 1}^{n} \int_{U_i} \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx.
\end{equation*}
We will now establish the following Poincare-type inequality, 
\begin{equation}
\label{pf:prop_isometry_3}
\int_{U_i} \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx \leq \frac{2^{d + 2}}{a_1(\Xset,d) \nu_d K(1/2)} \wt{\delta}^2 \wt{E}_{2\wt{\delta}}(f,U_i)
\end{equation}
which holds for all $i = 1,\ldots,n$. Summing up over $i$ implies
\begin{equation*}
\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)}^2 \leq \frac{2^{d+2}}{a_1(\Xset,d)K(1/2)p_{\min}} \cdot \wt{E}_{2\wt{\delta}}(f) \leq \frac{2^{d+3}}{a_1(\Xset,d)K(1/2)p_{\min}} A_1(\Xset,d,K,p_{\min}) D_2(f)
\end{equation*}
with the latter inequality following just as in the proof of Proposition~\ref{prop:dirichlet_energies}; the Lemma then follows by plugging this inequality into~\eqref{pf:prop_isometry_2}.

Now it remains to show~\eqref{pf:prop_isometry_3}. The key point is that for any pair of points $x,x' \in \Xset$ such that $\|x - x'\| \leq r$  the triangle inequality
\begin{equation*}
\bigl|f(x') - f(x)\bigr|^2 \leq 2\bigl|f(x') - f(z)\bigr|^2 + 2\bigl|f(z) - f(x)\bigr|^2
\end{equation*}
holds for each $z \in B\bigl(x,\wt{\delta}\bigr) \cap B\bigl(x',\wt{\delta}\bigr) \cap \Xset =: W$, and we have that
\begin{align*}
\bigl|f(x') - f(x)\bigr|^2 & \leq \frac{2}{\vol(W)}\biggl[\int_{W} \bigl|f(x') - f(z)\bigr|^2 \,dz + \int_{W} \bigl|f(z) - f(x)\bigr|^2 \,dz\biggr] \\
& \leq \frac{2}{\vol(W)K(1/2)}\biggl[\int_{\Xset} K\biggl(\frac{x' - z}{2\wt{\delta}}\biggr)\bigl|f(x') - f(z)\bigr|^2 \,dz + \int_{\Xset} K\biggl(\frac{z - x}{2\wt{\delta}}\biggr)\bigl|f(z) - f(x)\bigr|^2 \,dz\biggr] \\
& \leq \frac{2^{d + 1}}{a_1(\Xset,d) \nu_d \wt{\delta}^d K(1/2)}\biggl[\int_{\Xset} K\biggl(\frac{x' - z}{2\wt{\delta}}\biggr)\bigl|f(x') - f(z)\bigr|^2 \,dz + \int_{\Xset} K\biggl(\frac{z - x}{2\wt{\delta}}\biggr)\bigl|f(z) - f(x)\bigr|^2 \,dz\biggr]
\end{align*}
where the last inequality follows from Assumption~\ref{asmp:r_small_1}. We use this along with the Cauchy-Schwarz inequality to prove~\eqref{pf:prop_isometry_3}:
\begin{align*}
\int_{U_i} & \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx \\
 & =\int_{U_i} \biggl(n\cdot\int_{U_i} \bigl(f(x) - f(x')\bigr) \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx \\
& \leq n \cdot \int_{U_i} \int_{U_i} \bigl(f(x) - f(x')\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx \\
& \leq \frac{2^{d + 2}}{a_1(\Xset,d) \nu_d \wt{\delta}^d K(1/2)}n\int_{U_i} \int_{U_i} \int_{\Xset} K\biggl(\frac{x - z}{2\wt{\delta}}\biggr)\bigl|f(x) - f(z)\bigr|^2 \wt{\rho}_n(x') \wt{\rho}_n(x)  \,dx' \,dz \,dx \\
& \leq \frac{2^{d + 2}}{a_1(\Xset,d) \nu_d \wt{\delta}^d K(1/2) p_{\min}} \int_{U_i} \int_{U_i} K\biggl(\frac{x - z}{2\wt{\delta}}\biggr)\bigl|f(x) - f(z)\bigr|^2 \wt{\rho}_n(z) \wt{\rho}_n(x)  \,dz \,dx \\
& =  \frac{2^{d + 2}}{a_1(\Xset,d) \nu_d K(1/2) p_{\min}} \wt{\delta}^2 \wt{E}_{2\wt{\delta}}(f,U_i).
\end{align*}

\textit{Proof of (2).}
\mbox{}\\
\mbox{}\\

By the triangle inequality,
\begin{align}
\Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(P)}^2 - \|u\|_{\Leb^2(P_n)}^2\Bigr| & \leq \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(P)}^2 - \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2\Bigr| + \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 - \|u\|_{\Leb^2(P_n)}^2\Bigr| \nonumber \\
& \leq A_1(\theta + \wt{\delta}) \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 + \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 - \|u\|_{\Leb^2(P_n)}^2\Bigr| \nonumber\\
& = A_1(\theta + \wt{\delta}) \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 + \Bigl(\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} + \|u\|_{\Leb^2(P_n)}\Bigr) \cdot \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|u\|_{\Leb^2(P_n)}\Bigr| \label{pf:prop_isometry_4}
\end{align}
To upper bound the second term in the above expression, we first note that~$\|u\|_{\Leb^2(P_n)} = \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}$, and thus
\begin{align}
\Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|u\|_{\Leb^2(P_n)} \Bigr| & = \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}\Bigr| \nonumber \\
& \overset{(iii)}{\leq} \|\Lambda_{\wt{r}}\wt{\mc{P}}^{\star}u - \wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)} \nonumber \\
& \overset{(iv)}{\leq} r \sqrt{\frac{2}{\sigma_K p_{\min}} E_{\wt{r}}(\wt{\mc{P}}^{\star}u)} \nonumber \\
& \overset{(v)}{\leq} r \sqrt{\frac{2}{\sigma_K p_{\min}} \Bigl(1 + A_3(d,K,p)\frac{\wt{\delta}}{r}\Bigr) b_r(u)} \label{pf:prop_isometry_5}
\end{align}
where $(iii)$ follows by the triangle inequality, $(iv)$ follows from Lemma~\ref{lem:smoothening_error}, and $(v)$ follows from~\eqref{eqn:calder19_1} and Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}. On the other hand, by~\eqref{eqn:calder19_2} and Lemma~\ref{lem:smoothening_error},
\begin{align*}
\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 & \leq \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \|\wt{\mc{I}}u\|_{\Leb^2(P)}^2 \\
& \leq \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(P)}^2 \\
& \leq \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)^2 \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}^2 \\
& = \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)^2 \|u\|_{\Leb^2(P_n)}^2
\end{align*}
and plugging this estimate along with~\eqref{pf:prop_isometry_5} back into~\eqref{pf:prop_isometry_4}, we obtain part (2) of Proposition~\ref{prop:isometry}.

\paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}.} \mbox{} \\
\mbox{} \\
Recalling that $\bigl(\wt{P}f\bigr)(X_i) = n \cdot \int_{U_i} f(x) \wt{p}_n(x)$, the key point is that
\begin{equation*}
\biggl(\bigl(\wt{P}f\bigr)(X_i) - \bigl(\wt{P}f\bigr)(X_j)\biggr)^2 \leq n^2 \cdot \int_{U_i} \int_{U_j} \bigl(f(x') - f(x)\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx.
\end{equation*}
Additionally, the non-increasing and Lipschitz properties of $K$ imply that for any $x \in U_i$ and $x' \in U_j$, 
\begin{equation*}
K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) \leq K\biggl(\frac{\bigl(\|x' - x\| - 2\wt{\delta}\bigr)_{+}}{r}\biggr) \leq K\biggl(\frac{\|x' - x\|}{r + 2\wt{\delta}}\biggr) + \frac{2L_K\wt{\delta}}{r}\1\Bigl\{\|x' - x\| \leq r + 2\wt{\delta}\Bigr\}
\end{equation*}
Using this along with the Lipschitz property of $K$, we get
\begin{align*}
b_r(f) & = \frac{1}{n^2r^{d + 2}} \sum_{i,j = 1}^n \Bigl(\bigl(\wt{P}f\bigr)(X_i) - \bigl(\wt{P}f\bigr)(X_j)\Bigr)^2 K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) \\
& \leq \frac{1}{r^{d + 2}} \sum_{i,j = 1}^n \int_{U_i} \int_{U_j}  \bigl(f(x') - f(x)\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) \,dx' \,dx \\
& \leq \frac{1}{r^{d + 2}} \sum_{i,j = 1}^n \int_{U_i} \int_{U_j}  \bigl(f(x') - f(x)\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) \biggl[K\biggl(\frac{\|x' - x\|}{r + 2\wt{\delta}}\biggr) + \frac{2L_K\wt{\delta}}{r}\1\Bigl\{\|x' - x\| \leq r + 2\wt{\delta}\Bigr\}\biggr] \,dx' \,dx \\
& = \wt{E}_{r + 2\wt{\delta}}(f) + \frac{2L_K\wt{\delta}}{r}\wt{E}_{r + 2\wt{\delta}}(f; \1_{[0,1]})
\end{align*}
for $\1_{[0,1]}(t) = \1\{0 \leq t \leq 1\}$. But clearly $\wt{E}_{r + 2\wt\delta}(f; \1_{[0,1]}) \leq 1/(K(1/2))\wt{E}_{2r + 4\wt{\delta}}(f)$, and the Lemma is shown.

\paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected}.}
\mbox{}\\
\mbox{}\\
Suppose $f \in C^{\infty}(\wb{\Xset})$, which we may do without loss of generality due to the density of $C^{\infty}(\wb{\Xset})$ in  $\Leb^2(\Xset)$.

Taylor expanding $f(x')$ about $x' = x$, applying first Cauchy-Schwarz' and then Jensen's inequality, and then using the Lipschitz property of $p$, we have
\begin{align*}
E_r(f) & = \frac{1}{r^{d + 2}}\int_{\Xset} \int_{\Xset} (f(x') - f(x))^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \\
& = \frac{1}{r^{d + 2}}\int_{\Xset} \int_{\Xset} \biggl(\sum_{i = 1}^{d} (x' - x)^{e_i} \int_0^1 f^{(e_i)}\bigl(x + t(x' - x)\bigr) \,dt\biggr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \\
& \leq \frac{1}{r^{d + 2}}\int_{\Xset} \int_{\Xset} \|x' - x\|^2 \sum_{i = 1}^{d}\biggl(\int_0^1 f^{(e_i)}\bigl(x + t(x' - x)\bigr) \,dt\biggr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \\
& \leq \frac{1}{r^{d + 2}}\int_{\Xset} \int_{\Xset} \int_0^1   \|x' - x\|^2 \biggl\{\sum_{i = 1}^{d} \Bigl(f^{(e_i)}\bigl(x + t(x' - x)\bigr)\Bigr)^2\biggr\} K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x)   \,dt   \,dx' \,dx \\
& \leq \frac{(1 + L_pr)^2}{r^{d + 2}}\int_{\Xset} \int_{\Xset} \int_0^1  \|x' - x\|^2 \biggl\{\sum_{i = 1}^{d} \Bigl(f^{(e_i)}\bigl(x + t(x' - x)\bigr)\Bigr)^2\biggr\} K\biggl(\frac{\|x' - x\|}{r}\biggr) p\bigl(x + t(x' - x)\bigr)^2   \,dt   \,dx' \,dx.
\end{align*} 
Then letting $z = (x' - x)/r$ and $\wt{x} = x + trz$, we obtain
\begin{align*}
E_r(f) & \leq (1 + L_pr)^2 \int_{\Xset} \int_{\mathcal{Z}_r(x)} \int_{0}^{1} \|z\|^2 \biggl\{\sum_{i = 1}^{d} \Bigl(f^{(e_i)}\bigl(x + trz\bigr)\Bigr)^2\biggr\} K\bigl(\|z\|\bigr) p(x + trz)^2 \,dt \,dz \,dx \\
& = (1 + L_pr)^2 \int_{B(0,1)} \int_{0}^{1}  \int_{\Xset_r(z)} \|z\|^2 \biggl\{\sum_{i = 1}^{d} \Bigl(f^{(e_i)}\bigl(x + trz\bigr)\Bigr)^2\biggr\} K\bigl(\|z\|\bigr) p(x + trz)^2 \,dx \,dt \,dz \\
& \leq (1 + L_pr)^2 \int_{B(0,1)} \int_{0}^{1} \|z\|^2 K\bigl(\|z\|\bigr) \biggl\{\int_{\wt{\Xset}_r(t,z)} \|\nabla f(\wt{x})\|^2 p(\wt{x})^2 \,d\wt{x}\biggr\} \,dt \,dz  \\
& \leq (1 + L_pr)^2 \int_{B(0,1)} \int_{0}^{1} \|z\|^2 K\bigl(\|z\|\bigr) \biggl\{\int_{\Xset} \|\nabla f(\wt{x})\|^2 p(\wt{x})^2 \,d\wt{x}\biggr\} \,dt \,dz \\
& = (1 + L_pr)^2 \int_{B(0,1)} \|z\|^2 K\bigl(\|z\|\bigr) \,dz 
\end{align*}
where $\mathcal{Z}_r(x) = B(0,1) \cap (\Xset - x)/r$, $\Xset_r(z) = \Xset \cap( \Xset - zr)$, $\wt{\Xset}_r(t,z) = \Xset_r(z) + trz$, and we note that since
\begin{align*}
\wt{x} \in \wt{\Xset}_r(t,z) & \Longrightarrow \wt{x} = x + trz ~~ \textrm{for some $x \in \Xset - zr$ and $t \in [0,1]$} \\
& \Longrightarrow \wt{x} \in \Xset,~~ \textrm{(since $x \in \Xset$, $x + rz \in \Xset$ and $\Xset$ is connected)}
\end{align*}
it must be that $\wt{X}_r(t,z) \subseteq \Xset$ for all $z \in B(0,1)$ and $t \in [0,1]$.

\paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}.}
\mbox{} \\
\mbox{} \\
For any $a \in \Reals$, $\Lambda_rf$ satisfies the identity
\begin{equation*}
\Lambda_rf(x) = a + \frac{1}{r^d\tau(x)}\int_{\Xset} \eta_r(x',x)\bigl(f(x') - a\bigr)\,dx'
\end{equation*}
and by differentiating we obtain
\begin{equation*}
\bigl(\nabla \Lambda_rf\bigr)(x)= \frac{1}{r^d\tau(x)}\int_{\Xset} \bigl(\nabla \eta_r(x',\cdot)\bigr)(x)\bigl(f(x') - a\bigr)\,dx' + \nabla\bigl(\tau^{-1}\bigr)(x)\cdot \frac{1}{r^d}\int_{\Xset} \eta_r(x',x)\bigl(f(x') - a\bigr)\,dx'
\end{equation*} 
Plugging in $a = f(x)$, we get $\nabla\Lambda_rf(x) = J_1(x)/\tau(x) + J_2(x)$ for
\begin{equation*}
J_1(x) := \frac{1}{r^d}\int_{\Xset} \bigl(\nabla \eta_r(x',\cdot)\bigr)(x)\bigl(f(x') - f(x)\bigr)\,dx',~~ J_2(x) := \nabla\bigl(\tau^{-1}\bigr)(x)\cdot \frac{1}{r^d}\int_{\Xset} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)\,dx'.
\end{equation*}
To upper bound $\bigl\|J_1(x)\bigr\|^2$, we first compute the gradient of $\eta_r(x',\cdot)$,
\begin{align*}
\bigl(\nabla\eta_r(x',\cdot)\bigr)(x) & = \frac{1}{r} \psi'\biggl(\frac{\|x'  - x\|}{r}\biggr) \frac{(x - x')}{\|x' - x\|} \\
& = \frac{1}{\sigma_Kr^2} K\biggl(\frac{\|x' - x\|}{r}\biggr) (x' - x),
\end{align*}
so that by the Cauchy-Schwarz inequality,
\begin{align*}
\bigl\|J_1(x)\bigr\|^2 & = \frac{1}{\sigma_K^2 r^{4 + 2d}} \Biggl[\int_{\Xset} \bigl(f(x') - f(x)\bigr)K\biggl(\frac{\|x' - x\|}{r}\biggr)(x' - x)\Biggr]^2 \,dx' \\
& \leq \frac{1}{\sigma_K^2r^{4 + 2d}} \biggl[\int_{\Xset}K\biggl(\frac{\|x' - x\|}{r}\biggr)\|x' - x\|^2\,dx'\biggr] \biggl[\int_{\Xset}K\biggl(\frac{\|x' - x\|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2\,dx'\biggr] \\
& \leq \frac{1}{\sigma_K r^{2 + d}}\int_{\Xset}K\biggl(\frac{\|x' - x\|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2\,dx'.
\end{align*}
To upper bound $\bigl\|J_2(x)\bigr\|^2$, we use the Cauchy-Schwarz inequality along with the observation $\eta_r(x',x) \leq \frac{1}{\sigma_K} K\bigl(\|x' - x\|/r\bigr)$ to deduce
\begin{align*}
\bigl\|J_2(x)\bigr\|^2 & \leq \Bigl\|\nabla\bigl(\tau^{-1}\bigr)(x)\Bigr\|^2\frac{1}{r^{2d}} \biggl[\int_{\Xset}\eta_r(x',x) \,dx'\biggr] \cdot \biggl[\int_{\Xset} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr] \\
& = \Bigl\|\nabla\bigl(\tau^{-1}\bigr)(x)\Bigr\|^2\frac{\tau(x)}{r^d} \int_{\Xset} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)^2 \,dx'\\ 
& \leq \Bigl\|\nabla\bigl(\tau^{-1}\bigr)(x)\Bigr\|^2\frac{\tau(x)}{\sigma_K r^d}\int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx' \\
& \leq \frac{16}{\sigma_K^2r^{2 + d}} \biggl[\int_{\Xset} K\biggl(\frac{|x' - x|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr].
\end{align*}
where the last inequality follows by the estimates on $\tau$ and $\nabla \tau$ given in Lemma~\ref{lem:tau_bound}. Combining our bounds on $\bigl\|J_1(x)\bigr\|^2$ and $\bigl\|J_2(x)\bigr\|^2$ along with the lower bound on $\tau(x)$ in Lemma~\ref{lem:tau_bound} and integrating over $\Xset$, we have
\begin{align*}
\sigma_K D_2(\Lambda_r f) & = \sigma_K\int_{\Xset} \biggl\|\Bigl(\nabla \Lambda_rf)(x)\biggr\|^2 p^2(x) \,dx \\
& \leq 2 \sigma_K \int_{\Xset} \Biggl(\frac{\bigl\|J_1(x)\bigr\|^2}{\tau^2(x)} + \bigl\|J_2(x)\bigr\|^2\Biggr) p^2(x) \,dx \\
& \leq \frac{40}{\sigma_K r^{2 + d}} \int_{\Xset} \int_{\Xset} K\biggl(\frac{|x' - x|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 p^2(x) \,dx' \,dx \\
& \leq \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} E_r(f),
\end{align*}
completing the proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}.

\paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}.}
\mbox{} \\
\mbox{} \\
For brevity, we write $\wt{r} := r - 2\wt{\delta}$. We begin by expanding the energy $\wt{E}_{r - 2\wt{\delta}}\bigl(\wt{\mc{P}}^{\star}u\bigr)$ as a double sum of double integrals,
\begin{align*}
\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & = \frac{1}{\wt{r}^{d + 2}} \sum_{i = 1}^{n} \sum_{j = 1}^{n} \int_{U_i} \int_{U_j} \Bigl(u(X_i) - u(X_j)\Bigr)^2 K\biggl(\frac{\|x' - x\|}{\wt{r}}\biggr) \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx
\end{align*}
We next use the Lipschitz property of the kernel $K$---in particular that for $x \in U_i$ and $x' \in U_j$,
\begin{equation*}
K\biggl(\frac{\|x' - x\|}{\wt{r}}\biggr) \leq K\biggl(\frac{\|x_i - x_j\|}{r}\biggr) + \frac{2L_K\wt{\delta}}{\wt{r}} \cdot \1\biggl\{\frac{\|x' - x\|}{\wt{r}}\biggr\}
\end{equation*}
---to conclude that
\begin{align}
\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & \leq \frac{1}{n^2\wt{r}^{d + 2}} \sum_{i = 1}^{n} \sum_{j = 1}^{n} \Bigl(u(X_i) - U(X_j)\Bigr)^2 K\biggl(\frac{\|x_i - x_j\|}{r}\biggr) + \frac{2L_K\wt{\delta}}{\wt{r}}\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \nonumber \\
& \leq \biggl(1 + 4\bigl(2^{d + 2} - 1\bigr)\frac{\wt{\delta}}{r}\biggr) b_r(u) + \frac{2L_K\wt{\delta}}{\wt{r}}\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \label{pf:lem_first_order_graph_sobolev_seminorm_discretized_1}.
\end{align}
We focus now on the factor of $\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,1_{[0,1]})$. For any $x$ and $x' \in \Xset$ such that $\|x - x'\| \leq r$, note that by assumption $z_{xx'}:= \frac{1}{2}x + \frac{1}{2}x' \in \Xset \cap B(x,\wt{r}/2) \cap B(x',\wt{r}/2)$. Moreover by the triangle inequality
\begin{equation*}
\bigl|f(x) - f(x')\bigr|^2 \leq 2\bigl|f(x) - f(z_{xx'})\bigr|^2 + 2\bigl|f(z_{xx'}) - f(x')\bigr|^2
\end{equation*}
so that
\begin{align*}
\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) & \leq \int_{\Xset} \int_{\Xset \cap B(x,r)} \Bigl[ 2\bigl|f(x) - f(z_{xx'})\bigr|^2 +  2\bigl|f(z_{xx'}) - f(x')\bigr|^2\Bigr] \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx.
\end{align*}
Now, by a change of variables along with the Lipschitz property of $p$, we can deduce that
\begin{align*}
 \int_{\Xset} \int_{\Xset \cap B(x,r)} \bigl|f(x) - f(z_{xx'})\bigr|^2 \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx & \leq 2^d \bigl(1 + L_p\wt{r}\bigr)\int_{\Xset} \int_{\Xset \cap B(x,r/2)} \bigl|f(x) - f(z)\bigr|^2 \wt{p}_n(z) \wt{p}_n(x) \,dz \,dx \\
 & = 2^d \bigl(1 + L_p\wt{r}\bigr) \wt{E}_{\wt{r}/2}\bigl(f; \1_{[0,1]}\bigr),
\end{align*}
and therefore by symmetry
\begin{equation*}
\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \leq 2^{d + 2} \bigl(1 + L_p\wt{r}\bigr) \wt{E}_{\wt{r}/2}\bigl(f; \1_{[0,1]}\bigr).
\end{equation*}
Finally, noting that $\wt{E}_{\wt{r}/2}\bigl(f; \1_{[0,1]}) \leq \frac{1}{K(1/2)} \wt{E}_{\wt{r}}\bigl(f)$, the proof is completed by plugging our upper bound on $\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr)$ back into~\eqref{pf:lem_first_order_graph_sobolev_seminorm_discretized_1}.
\paragraph{Proof of Lemma~\ref{lem:tau_bound}.}
\mbox{}\\
\mbox{}\\
First prove the estimates of $\tau(x)$, and then upper bound on $\|\nabla\tau(x)\|$. Substituting $z = (x' - x)/r$ and using~\ref{asmp:r_small_3}, we have that
\begin{align*}
\tau(x) & = \frac{1}{r^d} \int_{\Xset} \psi\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx' \\
& =  \int_{\mc{B}_x} \psi\bigl(\|z\|\bigr) \,dz \\
& \geq a_1(\Xset,d) \cdot \int_{B(0,1)} \psi(\|z\|) \,dz
\end{align*}
We will now show that $\int_{B(0,1)} \psi(\|z\|) \,dz = 1$, which will imply our estimate of $\tau(x)$. To see this identity, note that on the one hand, by converting to polar coordinates and integrating by parts we obtain
\begin{align*}
\int_{B(0,1)} \psi\bigl(\|z\|\bigr) \,dz = d \nu_d \int_{0}^{1} \psi(t) t^{d - 1} \,dt = \nu_d \int_{0}^{1} \psi'(t) t^{d} \,dt = \frac{\nu_d}{\sigma_K} \int_{0}^{1} t^{d + 1} K(t) \,dt;
\end{align*}
on the other hand, again converting to polar coordinate, we have
\begin{equation*}
\sigma_K = \frac{1}{d} \int_{\Reals^d} \|x\|^2 K(\|x\|) \,dx = \nu_d \int_{0}^{1}t^{d + 1} K(t) \,dt.
\end{equation*}

Finally, we upper bound $\|\nabla\tau(x)\|^2$. Exchanging derivative and integral, we have
\begin{align*}
\nabla\tau(x) = \frac{1}{r^d} \int_{\Xset} \bigl(\nabla \eta_r(x',\cdot)\bigr)(x) \,dx' = \frac{1}{\sigma_K r^{d + 2}} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr)(x' - x)\,dx',
\end{align*}
and by the Cauchy-Schwarz inequality,
\begin{equation*}
\|\nabla\tau(x)\|^2 \leq \frac{1}{\sigma_K^2 r^{2d + 4}} \biggl[\int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr)\,dx'\biggr] \int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr)\|x' - x\|^2\,dx', \leq \frac{1}{\sigma_K r^{2}}
\end{equation*}
concluding the proof of Lemma~\ref{lem:tau_bound}. One note: while $\nabla\tau(x) = 0$ when $B(x,r) \in \Xset$, near the boundary the upper bound we derived by using Cauchy-Schwarz appears tight.

\paragraph{Proof of Lemma~\ref{lem:smoothening_error}.}
\mbox{}\\
\mbox{}\\
By Jensen's inequality and Lemma~\ref{lem:tau_bound},
\begin{align*}
\Bigl[\Lambda_rf(x)\Bigr]^2 & \leq \frac{1}{r^d \tau(x)}\int_{\Xset} \eta_r(x',x) \bigl[f(x')\bigr]^2 \,dx' \\
& \leq \frac{2}{r^d p_{\min}}\int_{\Xset} \eta_r(x',x) \bigl[f(x')\bigr]^2 p(x') \,dx'
\end{align*}
Then integrating over $x$ and swapping the order of integraton, we have
\begin{align*}
\bigl\|\Lambda_rf\bigr\|_{\Leb^2(P)}^2 & \leq \frac{2}{r^d p_{\min}} \int_{\Xset} \int_{\Xset} \eta_r(x',x) \bigl[f(x')\bigr]^2 p(x') p(x) \,dx' \,dx \\ 
& \leq \frac{2p_{\max}}{p_{\min}} \int_{\Xset} \bigl[f(x')\bigr]^2 p(x') \,dx' \\
& = \frac{2p_{\max}}{p_{\min}} \|f\|_{\Leb^2(P)}^2.
\end{align*}

Noting that $\Lambda_ra = a$ for any $a \in \Reals$, by the Cauchy-Schwarz inequality we have that for all $x \in \Xset$,
\begin{align*}
\bigl|\Lambda_rf(x) - f(x)\bigr|^2 & = \biggl[\frac{1}{r^d\tau(x)} \int_{\Xset} \eta_r(x',x) \bigl(f(x') - f(x)\bigr) \,dx'\biggr]^2 \\
& \leq \frac{1}{r^{2d} \tau^2(x)} \biggl[\int_{\Xset} \eta_r(x',x) \,dx'\biggr] \cdot \biggl[\int_{\Xset} \eta_r(x',x) \bigl(f(x') - f(x)\bigr)^2 \,dx'\biggr] \\
& = \frac{1}{r^d \tau(x)} \int_{\Xset} \eta_r(x',x) \bigl(f(x') - f(x)\bigr)^2 \,dx'. \\
& \leq \frac{1}{r^d \tau(x) p_{\min}} \int_{\Xset} \eta_r(x',x) \bigl(f(x') - f(x)\bigr)^2 p(x') \,dx'.
\end{align*}
Then integrating over $\Xset$ with respect to $p$ yields~\eqref{eqn:smoothening_error_energy}.

\clearpage

\section{Miscellaneous}

\paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb} (simplified setting).}
\mbox{} \\
\mbox{} \\
Assume $d = 1$ and $\Xset = [0,1]$. For any $a \in \Reals$,
\begin{equation*}
\Lambda_rf(x) = a + \frac{1}{r\tau(x)}\int_{0}^{1} \eta_r(x',x)\bigl(f(x') - a\bigr)\,dx'
\end{equation*}
Differentiating with respect to $x$, we obtain
\begin{equation*}
\frac{d}{dx}\Lambda_rf = \frac{1}{r\tau(x)}\int_{0}^{1} \frac{\partial}{\partial x}\eta_r(x',x)\bigl(f(x') - a\bigr)\,dx' + \frac{d}{dx}\bigl(\tau^{-1}(x)\bigr) \frac{1}{r}\int_{0}^{1} \eta_r(x',x)\bigl(f(x') - a\bigr)\,dx'
\end{equation*} 
and plugging in $a = f(x)$, we get $\frac{d}{dx}\Lambda_rf(x) = J_1(x)/\tau(x) + J_2(x)$ for
\begin{equation*}
J_1(x) := \frac{1}{r}\int_{0}^{1} \frac{\partial}{\partial x}\eta_r(x',x)\bigl(f(x') - f(x)\bigr)\,dx',~~ J_2(x) := \frac{d}{dx}\bigl(\tau^{-1}(x)\bigr)\frac{1}{r}\int_{0}^{1} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)\,dx'.
\end{equation*}
To upper bound $\bigl|J_1(x)\bigr|^2$, we first compute the partial derivative
\begin{align*}
\frac{\partial}{\partial x}\eta_r(x',x) & = \frac{1}{r} \psi'\biggl(\frac{|x'  - x|}{r}\biggr) \cdot \frac{(x - x')}{\|x' - x\|} \\
& = \frac{1}{\sigma_Kr^2} K\biggl(\frac{\|x' - x\|}{r}\biggr) \cdot (x' - x),
\end{align*}
so that by the Cauchy-Schwarz inequality,
\begin{align*}
\bigl|J_1(x)\bigr|^2 & = \frac{1}{\sigma_K^2 r^6} \Biggl[\int_{0}^{1} K\biggl(\frac{\|x' - x\|}{r}\biggr) \cdot (x' - x) \cdot \bigl(f(x') - f(x)\bigr)\Biggr]^2 \,dx' \\
& \leq \frac{1}{\sigma_K^2r^6} \biggl[\int_{0}^{1}K\biggl(\frac{\|x' - x\|}{r}\biggr)(x' - x)^2\,dx'\biggr] \cdot \biggl[\int_{0}^{1}K\biggl(\frac{\|x' - x\|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2\,dx'\biggr] \\
& \leq \frac{1}{\sigma_K r^3}\int_{0}^{1}K\biggl(\frac{\|x' - x\|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2\,dx'.
\end{align*}
To upper bound $\bigl|J_2(x)\bigr|^2$, we note that $\eta_r(x',x) \leq \frac{1}{\sigma_K} K_r(x',x)$. Then by the Cauchy-Schwarz inequality again,
\begin{align*}
\bigl|J_2(x)\bigr|^2 & \leq \biggl[\frac{d}{dx}\bigl(\tau^{-1}(x)\bigr)\biggr]^2\frac{1}{r^2} \biggl[\int_{0}^1\eta_r(x',x) \,dx'\biggr] \cdot \biggl[\int_{0}^{1} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr] \\
& \leq \biggl[\frac{d}{dx}\bigl(\tau^{-1}(x)\bigr)\biggr]^2\frac{\tau(x)}{r}\biggl[\int_{0}^{1} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr]\\ 
& \leq \biggl[\frac{d}{dx}\bigl(\tau^{-1}(x)\bigr)\biggr]^2\frac{\tau(x)}{\sigma_K r}\biggl[\int_{0}^{1} K\biggl(\frac{|x' - x|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr] \\
& \leq \frac{8}{\sigma_Kr^3} \biggl[\int_{0}^{1} K\biggl(\frac{|x' - x|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr].
\end{align*}
where the last inequality follows by Lemma~\ref{lem:tau_bound}. Combining our bounds on $\bigl|J_1(x)\bigr|^2$ and $\bigl|J_2(x)\bigr|^2$, we have
\begin{align*}
\sigma_K D_2(\Lambda_r f) & = \sigma_K\int_{0}^{1} \biggl|\frac{d}{dx} \Lambda_rf(x)\biggr|^2 p^2(x) \,dx \\
& \leq \frac{18}{r^3} \int_{0}^{1} \int_{0}^{1} K\biggl(\frac{|x' - x|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 p^2(x) \,dx' \,dx \\
& = 18 E_r(f),
\end{align*}
completing the proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}. 
\paragraph{Proof of Lemma~\ref{lem:tau_bound} (simplified setting).} 
\mbox{} \\
\mbox{} \\
Assume $d = 1$ and $\Xset = [0,1]$. First we show the lower bound on $\tau(x)$, then the upper bound on $\tau(x)$, and finally the upper bound on $|\tau'(x)|$. Substituting $z = (x' - x)/r$ and integrating by parts, we have
\begin{align*}
\tau(x) & = \frac{1}{r} \int_{0}^{1} \psi\biggl(\frac{|x' - x|}{r}\biggr) \,dx' \\
& =  \int_{-x/r}^{(1 - x)/r}  \psi\bigl(\abs{z}\bigr) \,dz \\
& = z \psi\bigl(|z|\bigr) \bigg|_{-x/r}^{(1 - x)/r} - \int_{-x/r}^{(1 - x)/r}  z \psi'\bigl(\abs{z}\bigr) \,dz \\
& \geq \frac{1}{\sigma_K}\int_{-x/r}^{(1 - x)/r}  z^2 K\bigl(|z|\bigr) \,dz. \\
\end{align*}
By our smallness assumption on $r$, one of $x > r$ or $(1 - x) > r$ must be true; we will assume without loss of generality that $(1 - x) > r$. In that case,
\begin{equation*}
\int_{-x/r}^{(1 - x)/r}  z^2 K\bigl(|z|\bigr) \,dz \geq \int_{0}^{1} z^2K\bigl(|z|\bigr) = \frac{1}{2} \int_{-1}^{1} z^2K\bigl(|z|\bigr) \,dz = \frac{1}{2}\sigma_K,
\end{equation*}
so that $\tau(x) \geq 1/2$.

The upper bound on $\tau(x)$ follows along similar lines. Again substituting $z = (x' - x)/r$ and integrating by parts, we have
\begin{align*}
\tau(x) & = \frac{1}{r} \int_{0}^{1} \psi\biggl(\frac{|x' - x|}{r}\biggr) \,dx' \\
& \leq \int_{-\infty}^{\infty}  \psi\bigl(\abs{z}\bigr) \,dz \\
& = \frac{1}{\sigma_K}\int_{-\infty}^{\infty} z^2 K(z) \,dz = 1.
\end{align*}

Finally, we upper bound $|\tau'(x)|$:
\begin{align*}
\abs{\tau'(x)} & = \frac{1}{r} \int_{0}^{1} \frac{\partial}{\partial x} \eta_r(x',x) \,dx' \\
& = \frac{1}{\sigma_Kr^{3}} \int_{0}^{1} K\biggl(\frac{\|x' - x\|}{r}\biggr) \cdot (x' - x) \,dx' \\
& = \frac{1}{\sigma_Kr} \int_{-x/r}^{(1 - x)/r} z K\bigl(|z|\bigr) \,dz \\
& \leq \frac{1}{\sigma_Kr} \biggl[\int_{-\infty}^{\infty} K\bigl(|z|\bigr) \,dz\biggr] \cdot \biggl[\int_{-\infty}^{\infty}z^2 K\bigl(|z|\bigr) \,dz \biggr] \\
& \leq \frac{1}{r}.
\end{align*}
The first upper bound in the previous display follows from the Cauchy-Schwarz inequality. While much tighter upper bounds are possible when $B(x,r) \subseteq \Xset$---indeed $\tau'(x) = 0$ for such $x$---near the boundary this inequality is tight.

\clearpage

\bibliographystyle{plainnat}
\bibliography{../../graph_testing_bibliography}

\end{document}

### 2020-10-06 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing2.tex
Updated draft of our AISTATS submission.

@@
 \title{Minimax optimal Laplacian smoothing}
-\author{Alden Green}
+\author{Alden Green, Sivaraman Balakrishnan, and Ryan Tibshirani}
 \date{\today}
@@
 
-\textbf{Disclaimer:} This largely follows the same setup as in the draft I have written which combines Laplacian smoothing and Laplacian eigenmaps results.
-That draft --- document name \textit{graph\_regression2} --- should for the moment be considered obsolete.
+\begin{abstract}
+	\textcolor{red}{(TODO)}:
+\end{abstract}
 
@@
 
-\textbf{(1)} Laplacian smoothing is a graph-based approach to nonparametric regression.
-
-\begin{itemize}
-	\item In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Reals^d$, and
+In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Reals^d$, and
 \begin{equation}
@@
 \end{equation}
-	with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise.
-We will assume throughout that $P$ admits a density $p$ with respect to the volume form of $\mc{X}$ which is bounded away from $0$ and $\infty$, i.e.
+with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise. We will assume the following throughout:
+\begin{enumerate}[label=(P\arabic*)]
+	\item
+	\label{asmp:bounded_lipschitz_density} 
+	$P$ admits a density $p$ bounded away from $0$ and $\infty$, i.e.
 	\begin{equation*}
@@
 	\end{equation*}
+	Additionally, $p$ is Lipschitz, with Lipschitz constant $L_p$.
+\end{enumerate}
 Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present.
-	\item In graph-based nonparametric regression, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$.
+
+In graph-based nonparametric regression, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$.
 The neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$.
@@
 \label{eqn:neighborhood_graph}
-	{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr).
+{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}}{r}\Biggr).
 \end{equation*}
+where $\norm{\cdot} = \norm{\cdot}_{\Rd}$ is the Euclidean norm.
 Then the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$, and the graph Laplacian can be written as
@@
 \end{equation}
-	\item The Laplacian smoothing estimator $\wh{f}_{\LS}$ \citep{smola2003} is a penalized least squares estimator, given by
+
+The Laplacian smoothing estimator $\wh{f}_{\LS}$ \citep{smola2003} is a penalized least squares estimator, given by
 \begin{equation}
@@
 \begin{equation}
-	T_{\LS} = \frac{1}{n}\Bigl\|\wt{f}_{\LS}\Bigr\|_2^2 \label{eqn:laplacian_smoothing_test}
+T_{\LS} = \frac{1}{n}\bigl\|\wt{f}_{\LS}\bigr\|_2^2 \label{eqn:laplacian_smoothing_test}
 \end{equation}
 is in turn a reasonable statistic to assess whether or not $f_0 = 0$.
-\end{itemize}
 
-\textbf{(1.a) Motivation.} Laplacian smoothing has many advantages compared to other nonparametric regression methods. For instance:
-
+Laplacian smoothing has many advantages compared to other nonparametric regression methods. For instance:
 \begin{itemize}
@@
 
-\textbf{(1.b) Related work.} For this reason a considerable body of work has emerged analyzing the convergence of the graph Laplacian $\Lap_{{n,r}}$ to the continuum Laplace operator $\Delta_P: C^2(\Xset) \to \Leb^2(\Xset)$,
+\paragraph{Related work.} For this reason a considerable body of work has emerged analyzing the properties of Laplacian smoothing, as well as of graph-based methods more generally.
+Roughly speaking this work can divided into two categories.
+
+Works in the first category view the responses $Y = \theta + \varepsilon$ as noisy samples of a regression function $\theta \in \Reals^n$ defined on the vertices of \emph{fixed graph} $G$.
+and bound the statistical error as a function of $\theta$ and $G$.
+(In the estimation literature, see for instance \citep{wang2016, sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}; in the testing literature, see \citep{sharpnack2013,sharpnack2013b,sharpnack2015}.) Note that in our setting, the graph $G_{n,r}$ and the evaluations $(f_0(X_1),\ldots,f_0(X_n))$ are random rather than fixed, and our assumptions are placed only on the underlying objects $P$ and $f_0$ from which these random object arise.
+As such, both our analysis and conclusions are naturally quite different.
+
+Works in the second category prove the consistency of neighborhood graph Laplacian smoothing and other neighborhood graph-based methods in the large-sample limit.
+They do so by establishing the convergence of the graph Laplacian $\Lap_{{n,r}}$ to the continuum Laplace operator $\Delta_P: C^2(\Xset) \to \Leb^2(\Xset)$,
 \begin{equation}
@@
 \end{equation}
-as $n \to \infty$ and $r(n) \to 0$. This is meant in various senses:
+as $n \to \infty$ and $r \to 0$. This is meant in various senses:
 \begin{itemize}
-	\item Pointwise convergence, meaning $\Lap_{n,r}f \to \Delta_Pf$. 
-	\item Spectral convergence, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$.
-	\item Convergence of norms, meaning $f^T \Lap_{n,r} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$. 
+	\item Pointwise convergence, meaning that $\Lap_{n,r}f \to \Delta_Pf$ in an appropriate norm \citep{belkin03,belkin05,lafon04,hein05,singer06,gine06}. 
+	\item Spectral convergence, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$, where the latter are the Dirichlet eigenvalues of $\Delta_P$ \citep{belkin07,burago2014,trillos2018,trillos2019,calder2019}.\footnote{\label{footnote:dirichlet_eigenvalues} Formally, Dirichlet eigenvalues are solutions to the problem
+		$$
+		\Delta_Pg = \lambda g,\\
+		\restr{g}{\partial \Xset} = 0
+		$$} 
+	\item Convergence of norms, meaning $f^T \Lap_{n,r} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$ \citep{bousquet03,hein06,zhou11}. 
 \end{itemize}
-However, the statistical optimality of Laplacian smoothing is still not well understood.
+While these results imply \emph{consistency} of many different graph Laplacian based procedures, the statistical \emph{optimality} of Laplacian smoothing over neighborhood graphs is still not well understood.
 
-\textbf{(1.c) Our contributions.} Our main contributions lie in this latter direction, and they can all be summarized as follows: Laplacian smoothing methods are minimax optimal over Sobolev spaces.
-In more detail, we will show that when $f_0$ belongs to the Sobolev ball $H^1(\Xset;M)$:
+\paragraph{Our contributions.} Our three main contributions lie in this latter direction, and they can all be summarized as follows: Laplacian smoothing methods are minimax optimal over first order Sobolev spaces.
+In more detail, we will show that when $f_0$ belongs to the first order Sobolev ball $H^1(\Xset;M)$:
 \begin{itemize}
-	\item \textbf{Minimax optimal estimation.} With high probability, $\frac{1}{n}\norm{\wt{f}_{\LS} - f_0}_{2}^2 \lesssim n^{-2/(2 + d)}$.
+	\item \textbf{Minimax optimal estimation.} With high probability, $\frac{1}{n}\bigl\|\wt{f}_{\LS} - f_0\bigr\|_{2}^2 \lesssim M^{d/(2+d)}n^{-2/(2 + d)}$.
 	\item \textbf{Minimax optimal testing.}
-	A level-$\alpha$ test constructed using $T_{\LS}$ has non-trivial power whenever $\norm{f_0}_{\Leb^2(\Xset)}^2 \gtrsim n^{-4/(4 + d)}$. 
+	A level-$\alpha$ test constructed using $T_{\LS}$ has non-trivial power whenever $\norm{f_0}_{\Leb^2(\Xset)}^2 \gtrsim M^{2d/(4 + d)}n^{-4/(4 + d)}$. 
 	\item \textbf{Manifold adaptivity.}
@@
 \end{itemize}
-Finally, we will also establish that the Laplacian smoothing estimator can be altered to take advantage of higher-order smoothness assumptions on $f_0$.
-This last conclusion will hold only under very limited circumstances.
 
-\textbf{(1.d) Organization.}
+\paragraph{Organization.} We now describe the structure of the rest of this paper, which builds to the aforementioned minimax results.
+\begin{itemize}
+	\item In Section~\ref{sec:minimax_optimal_regression_sobolev_spaces}, for completeness we review some standard results regarding minimax nonparametric regression over Sobolev spaces.
+	\item In Section~\ref{sec:graph_sobolev_classes}, we introduce the graph Sobolev class $H^1(G_{n,r})$, which we view as a graph-based analog to the continuum Sobolev class $H^1(\Xset)$.
+We show that key properties of $H^1(\Xset)$ used to derive minimax rates are shared by $H^1(G_{n,r})$.
+	\item In Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we leverage the results of Section~\ref{sec:graph_sobolev_classes} to derive tight upper bounds on the error of Laplacian smoothing methods.
+	\item In Section~\ref{sec:simulations} we provide some experiments supporting our theory, and conclude in Section~\ref{sec:discussion} with some potential methodological and theoretical extensions.
+\end{itemize}
 
-\textcolor{red}{(TODO)}
+\textbf{Notation.}
 
-\textbf{(1.e) Notation.}
-
-\textcolor{red}{(TODO)}
+We use $A,A_1, \ldots$ and $C,C_1, \ldots$ to refer to constants which do not depend on $n$, $r$, or $f_0$; likewise with $a, a_1,\ldots$ and $c, c_1,\ldots$.
 
 \section{Minimax-optimal regression over Sobolev spaces}
+\label{sec:minimax_optimal_regression_sobolev_spaces}
 
-\textbf{(2)} Before we get to our main results, we briefly review minimax optimal estimation and testing rates over Sobolev classes.
+Before we get to our main results, we briefly review minimax optimal estimation and testing rates over Sobolev classes.
 We also single out a notable method which achieves these rates: \emph{smoothing splines}.
@@
 
-\paragraph{(2.a) Sobolev spaces.} We start by briefly introducing Sobolev spaces, emphasizing only the material relevant to our statistical context. 
+\paragraph{Sobolev spaces, and the Sobolev embedding theorem.} We start by briefly introducing Sobolev spaces, emphasizing only the material relevant to our statistical context.
+
+Roughly speaking, (Hilbert-)Sobolev spaces contain functions $f \in \Leb^2(\Xset)$ with derivatives which themselves belong to $\Leb^2(\Xset)$.
+More formally, for a given integer $s$, the Sobolev space $H^{s}(\Xset)$ consists of all functions $f \in \Leb^2(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ exists and belongs to $\Leb^2(\Xset)$.
+For such functions, the Sobolev seminorm $\seminorm{f}_{H^{s}(\Xset)}$ and norm $\norm{f}_{H^{s}(\Xset)}$ are given by
+\begin{equation*}
+\seminorm{f}_{H^s(\Xset)}^2 = \sum_{\abs{\alpha} = s}\int_{\mathcal{X}} \Bigl|\bigl(D^{\alpha}f\bigr)(x)\Bigr|^2 \,dx, ~~ \norm{f}_{H^{s}(\Xset)}^2 = \sum_{k = 0}^{s} \seminorm{f}_{H^k(\Xset)}^2
+\end{equation*}
+and for a given $M > 0$, the order $s$ Sobolev ball is $H^s(\Xset, M) = \set{f: \norm{f}_{H^s(\Xset)} \leq M}$.
+The graph Sobolev seminorm $f^T \Lap_{n,r} f$ should be viewed as a random, discrete approximation to the continuum first-order Sobolev seminorm $|f|_{H^1(\Xset)}^2$, and so in our analysis of Laplacian smoothing we will mainly focus on the first-order Sobolev space $H^1(\Xset)$ hereafter.
+
+Much of statistical analysis of nonparametric regression over Sobolev spaces relies on the Sobolev embedding theorem, which establishes that functions in Sobolev spaces also satisfy other, (even more) classical notions of smoothness.
+Concretely, the Sobolev space $H^1(\Xset)$ will continuously embed either into a Holder space $C^{0,q}(\Xset)$ or an $\Leb^p$ space, depending on the dimension $d$.
 \begin{itemize}
-	\item Roughly speaking, Sobolev spaces contain functions $f \in \Leb^p(\Xset)$ with derivatives which themselves belong to $\Leb^p(\Xset)$.
-More formally, for given integers $s$ and $p > 0$, the Sobolev space $W^{s,p}(\Xset)$ consists of all functions $f \in \Leb^p(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ exists and belongs to $\Leb^p(\Xset)$.
-For such functions, the $(s,p)$-Sobolev seminorm $\seminorm{f}_{W^{s,p}(\Xset)}$ and norm $\norm{f}_{W^{s,p}(\Xset)}$ are given by
-	\begin{equation*}
-	\seminorm{f}_{W^{s,p}(\Xset)}^p = \sum_{\abs{\alpha} = s}\int_{\mathcal{X}} \Bigl|\bigl(D^{\alpha}f\bigr)(x)\Bigr|^p \,dx, ~~ \norm{f}_{W^{s,p}(\Xset)}^p = \sum_{k = 0}^{s} \seminorm{f}_{W^{k,p}(\Xset)}^p
-	\end{equation*}
-	and for a given $M > 0$, the $(s,p)$-Sobolev ball is $W^{s,p}(\Xset, M) = \set{f: \norm{f}_{W^{s,p}(\Xset)} \leq M}$.
-In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space, and we adopt the usual convention of writing $H^s(\Xset) = W^{s,2}(\Xset)$ and $H^s(\Xset,M) = W^{s,2}(\Xset,M)$.
-Linear estimators such as $\wh{f}_{\LS}$\footnote{To be explicit, estimators of the form $\wh{f} = HY$ for some matrix $H \in \Reals^{n \times n}$.} can be very accurate over the Hilbert-Sobolev spaces $H^s(\Xset)$, and we will confine our attention to this case hereafter.
-	\item \textcolor{red}{(TODO)}: Introduce the Sobolev embedding theorem, for at least three reasons: (1) to facilitate later discussion of why smoothing splines are ill-posed when $2s < d$; (2) to make explicit why the statistical problem isn't ill-posed when $2s < d$ in the random design setting; (3) to make clear why the testing problem is hopeless when $4s < d$ without imposing additional conditions.
+	\item When $d = 1$, any function $f \in H^1(\Xset)$ satisfies $\|f\|_{H^1(\Xset)} \leq C\|f\|_{C^{0,q}(\Xset)}$ for $q = 1/2.$
+	\item When $d > 2$, any function $f \in H^1(\Xset)$ satisfies $\|f\|_{H^1(\Xset)} \leq C\|f\|_{\Leb^{p}(\Xset)}$ for $p = 2d/(d - 2)$. 
 \end{itemize}
+In the above, the exponents $p$ and $q$ cannot be improved upon.
 
-\paragraph{(2.b) Connecting Laplacian smoothing to smoothing splines.} 
-To better motivate Laplacian smoothing as a tool well-suited for nonparametric regression, we review their connection to one of the most classical methods of nonparametric regression: smoothing splines.
-\begin{itemize}
-	\item Smoothing splines are particularly well-suited for nonparametric regression over Sobolev spaces.
-This is because smoothing splines are penalized least squares estimators where the penalty is itself the Sobolev semi-norm.
+When $d = 1$, the embedding of $H^1(\Xset)$ into $C^{0,q}(\Xset)$ facilitates a general analysis of (penalized) least estimators over $H^1(\Xset)$, using arguments from empirical process theory.
+However when $d \geq 2$, such arguments fall apart: for example, good bounds on the metric entropy of $H^1(\Xset)$ are not known in this setting.
+More troubling still, least squares estimators over $H^1(\Xset)$ are not even well-posed when $d \geq 2$.
+
+For these reasons, analysis of least squares estimators for the nonparametric regression problem over Sobolev spaces typically assume $d = 1$---or more generally that $f_0 \in H^s(\Xset)$ for some $2s > d$.
+However, the Laplacian smoother $\wh{f}_{\LS}$ is a special least-squares estimator: it has a nice closed-form solution amenable to a bias-variance decomposition.
+We therefore do not need to rely on empirical process theory, and so we will be able to analyze $\wh{f}_{\LS}$ and $T_{\LS}$ beyond the univariate setting.
+
+\paragraph{Connecting Laplacian smoothing to smoothing splines.} 
+To better motivate Laplacian smoothing as a tool for nonparametric regression, we review its connection to one of the most classical methods of nonparametric regression: smoothing splines.
+
+Smoothing splines are specially designed for regression over Sobolev spaces: they are penalized least squares estimators where the penalty is itself the Sobolev semi-norm.
 Mathematically, the smoothing spline estimator $\wh{f}_{\SM}$ is defined as the solution to the following optimization problem,
@@
 \label{eqn:smoothing_spline}
-	\wt{f} = \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \cdot \seminorm{f}_{H^{1}(\Xset)}^2\biggr\}.
+\wt{f}_{\SM} = \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \cdot \seminorm{f}_{H^{1}(\Xset)}^2\biggr\}.
 \end{equation}
-	In~\eqref{eqn:smoothing_spline}, the minimum can either be taken over all $f \in H^1(\Xset)$ or over all $f \in C^1(\Xset)$; we shall refer to the former as smoothing splines, and the latter as \emph{Holder} smoothing splines.
+In~\eqref{eqn:smoothing_spline}, the minimum can either be taken over all $f \in H^1(\Xset)$ or over all $f \in C^1(\Xset)$; we shall refer to the former as smoothing splines, and the latter as \emph{Lipschitz} smoothing splines.
 
-	\textcolor{red}{(TODO)}: Check the classical spline references, and/or Ryan, for data on (1) what the canonical domain of minimization is for smoothing splines, and (2) whether the representer theorem depends holds for either domain of minimization.
-
-	\item By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can tie the penalty used by the smoothing spline estimator $\wh{f}_{\SM}$ to that used by the Laplacian smoothing estimator $\wh{f}_{\LS}$.
+By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can tie the penalty used by the smoothing spline estimator $\wh{f}_{\SM}$ to that used by the Laplacian smoothing estimator $\wh{f}_{\LS}$.
 Suppose for the moment that $P$ is the uniform distribution over $\Xset$.
-Using the definition~\eqref{eqn:laplace_operator}, integrating by parts, and invoking the aforementioned convergence properties of $\Lap_{n,r}$, we have that for any $f \in C^{2}(\Xset)$ which satisfies appropriate boundary conditions,
+Using the definition~\eqref{eqn:laplace_operator}, integrating by parts, and invoking the aforementioned convergence properties of $\Lap_{n,r}$, we have that for any $f \in C^{2}(\Xset)$ which satisfies Dirichlet boundary conditions (see footnote~\ref{footnote:dirichlet_eigenvalues}),
 \begin{equation}
 \label{eqn:seminorm_convergence}
-	\abs{f}_{H^1(\Xset)} = \int_{\Xset} \bigl(\Delta_P f\bigr)(x) \cdot f(x) \,dx = \lim_{n \to \infty} \frac{1}{n^{2}r^{(d + 2)}}f^T \Lap_{n,r} f.
+\abs{f}_{H^1(\Xset)}^2 = \int_{\Xset} \bigl(\Delta_P f\bigr)(x) \cdot f(x) \,dx = \lim_{n \to \infty} \frac{1}{n^{2}r^{(d + 2)}}f^T \Lap_{n,r} f.
 \end{equation}
-	\textcolor{red}{(TODO)}: Formalize the boundary conditions---I believe either Neumann or Dirichlet boundary conditions should suffice?
-	\item \eqref{eqn:seminorm_convergence} establishes asymptotic equality (up to a scaling factor) between the penalties in~\eqref{eqn:smoothing_spline} and~\eqref{eqn:laplacian_smoothing}, when the distribution $P$ is uniform.
-But it also sheds light on additional feature of Laplacian smoothing, relevant when $P$ is non-uniform: the penalty $f^T \Lap_{n,r} f$ automatically adapts to the design density $p$, in a way that the penalty $|f|_{H^1(\Xset)}$ does not.
-This automatic adaptivity to $P$ turns out to be a formal advantage of Laplacian smoothing, which we will comment on further in Section~\ref{sec:minimax_optimal_laplacian_smoothing}.
-\end{itemize}
 
-\paragraph{(2.c) Minimax-optimality of smoothing splines.} 
-One reason all of this is intriguing is that, as mentioned previously, the smoothing spline estimator $\wh{f}_{\SM}$---and the resultant test statistic $T_{\SM} := \|\wt{f}_{\SM}\|_n^2$--- have strong statistical properties.
+\eqref{eqn:seminorm_convergence} establishes asymptotic equality (up to a scaling factor) between the penalties in~\eqref{eqn:smoothing_spline} and~\eqref{eqn:laplacian_smoothing}, when the distribution $P$ is uniform.
+But it also sheds light on an additional feature of Laplacian smoothing, relevant when $P$ is non-uniform: the penalty $f^T \Lap_{n,r} f$ automatically adapts to the design density $p$, in a way that the penalty $|f|_{H^1(\Xset)}^2$ does not.
+This automatic adaptivity to $P$ turns out to be a formal advantage of Laplacian smoothing relative to smoothing splines, as we will see in Section~\ref{sec:minimax_optimal_laplacian_smoothing}.
+
+\paragraph{Minimax-optimality of smoothing splines.} 
+One reason all of this is intriguing is that, as mentioned previously, the smoothing spline estimator $\wt{f}_{\SM}$---and the resultant test statistic $T_{\SM} := \|\wt{f}_{\SM}\|_n^2$--- have strong statistical properties.
 To be precise, they achieve the well-known minimax rates over Sobolev spaces.
-\begin{itemize}
-	\item  The minimax estimation rate over Sobolev balls, with error measured in $\Leb^2(\Xset)$ norm, is 
+
+Under~\ref{asmp:domain}, the minimax estimation rate over Sobolev balls is (see e.g. \citep{tsybakov2008_book})
 \begin{equation}
 \label{eqn:sobolev_space_estimation_minimax_rate}
-	\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{d/(2 + d)}n^{-2/(2 + d)}~~\textrm{for all $d \geq 1$.}
+\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{2d/(2 + d)}n^{-2/(2 + d)}
 \end{equation}
-
-	\textcolor{red}{(TODO)}: (1) You need a reference for this statement, and you already know one will be tricky to find because of ill-definedness of point evaluation.
-(2) This statement only holds under some condition on the density $p$.
-
-	\item On the other hand, the minimax critical radius over Sobolev balls---that is, the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has non-trivial power over all $H_{\epsilon}^1(\Xset,M) := H^1(\Xset,M) \cap \{f: \norm{f}_{\Leb^2(\Xset)} \geq \epsilon\}$---is much smaller,
+In the goodness-of-fit testing problem, we ask for a test function which can distinguish between the null and alternative hypotheses
+\begin{equation}
+\mathbf{H}_0: f_0 = 0, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{0\}
+\end{equation} 
+where $\mc{F}$ is some class of functions.
+The minimax critical radius is the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{F}_{\epsilon} := \mc{F} \cap \{f: \|f\|_{\Leb^2(\Xset)} \geq \epsilon\}$.
+Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing critical radius over $H^1(\Xset,M)$ is much smaller than the minimax estimation rate:
 \begin{equation}
 \label{eqn:sobolev_space_testing_critical_radius}
-	\epsilon_n^{\star}\Bigl(H^1(\Xset,M)\Bigr):= \inf\Biggl\{\epsilon: \inf_{\phi} \biggl[\Ebb_0[\phi] +  \sup_{f_0 \in H_{\epsilon}^1(\Xset,M)} \Ebb_{f_0}[1 - \phi]\biggr] \leq 2\alpha \Biggr\} \asymp n^{-2/(4 + d)}~~\textrm{for $1 \leq d < 4$,}
+\epsilon^2\Bigl(H^1(\Xset,M)\Bigr):= \inf\Biggl\{\epsilon^2: \inf_{\phi} \biggl[\Ebb_0[\phi] +  \sup_{f_0 \in H_{\epsilon}^1(\Xset,M)} \Ebb_{f_0}[1 - \phi]\biggr] \leq 2\alpha \Biggr\} \asymp M^{2d/(4 + d)}n^{-4/(4 + d)}~~\textrm{for $1 \leq d < 4$;}
 \end{equation}
+see \cite{ingster09} for a proof of this statement, and \cite{ariascastro2018} for a more extended discussion of the minimax paradigm in goodness-of-fit testing.
+When $d \geq 4$ we no longer have a continuous embedding of $H^1(\Xset)$ into $\Leb^4(\Xset)$; the functions in $H^1(\Xset)$ are very irregular, and the minimax rates in this regime are unknown.
 
-	\textcolor{red}{(TODO)}: (1) You need a reference for this statement (Ingster), and (2) you need to explain what happens when $d \geq 4$.
+The estimator $\wh{f}_{\SM}$ and the level-$\alpha$ test $\phi_{\SM} := \1\{T_{\SM} \geq {t}_{\alpha}\}$ are known to be minimax rate-optimal---meaning they achieve the rates in~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius}, respectively---when $d = 1$ (see e.g.
+\cite{vandergeer2000} for the estimation result, and \cite{liu2019} for the testing result).
+When $d > 1$, as a consequence of the Sobolev Embedding Theorem\footnote{Or to be more precise, as a consequence of the \emph{tightness} of the Sobolev Embedding Theorem.} the variational problem~\eqref{eqn:smoothing_spline} is not well-posed: as illustrated in e.g.
+\cite{green93}, the criterion is minimized by interpolating the responses $\mathbf{Y}$, and the resulting estimator $\wt{f}_{\SM}$ will be inconsistent.
+By replacing the domain of minimization $H^1(\Xset)$ by $C^1(\Xset)$, we can force the solution to~\eqref{eqn:smoothing_spline} to be less wild, and \cite{birge1993} show that the resulting Lipschitz smoothing spline is not merely consistent but in fact minimax optimal when $d = 2$.\footnote{At least in the estimation problem.
+We could not find either a positive or negative result for the testing problem.} However when $d > 2$, the Holder space $C^1(\Xset)$ is itself too large, and even the Lipschitz smoothing spline has risk provably larger than optimal \cite{birge1993}.
+Previewing things a bit, we will see in Section~\ref{sec:minimax_optimal_laplacian_smoothing} that a similar phenomenon occurs for the Laplacian smoother $\wh{f}$, but interestingly the breakdown point will come when $d > 4$, rather than $d > 1$ or $d > 2$.
 
-	\item The estimator $\wh{f}_{\SM}$ and the level-$\alpha$ test $\phi_{\SM} := \1\{T_{\SM} \geq {t}_{\alpha}\}$ are known to be minimax rate-optimal---meaning they achieve the rates in~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius}, respectively---when $d = 1$.
-When $d > 1$, as a consequence of the Sobolev Embedding Theorem\footnote{Or to be more precise, as a consequence of the \emph{tightness} of the Sobolev Embedding Theorem.} the variational problem~\eqref{eqn:smoothing_spline} is not well-posed: the criterion is minimized by interpolating the responses $\mathbf{Y}$, and the resulting estimator $\wt{f}$ will be inconsistent.
-By replacing the domain of minimization $H^1(\Xset)$ by $C^1(\Xset)$, we can force the solution to~\eqref{eqn:smoothing_spline} to be less wild, and the resulting Holder smoothing spline is not merely consistent but in fact minimax optimal when $d = 2$.\footnote{At least in the estimation problem.
-We could not find either a positive or negative result for the testing problem.} However when $d > 2$, the Holder space $C^1(\Xset)$ is itself too large, and even the Holder smoothing spline has risk provably larger than optimal.
-Previewing things a bit, we will see in Section~\ref{sec:minimax_optimal_laplacian_smoothing} that a similar phenomenon occurs for the Laplacian smoother $\wh{f}$; interestingly, the breakdown point will come when $d > 4$, rather than $d > 1$ or $d > 2$.
+Finally, it is well known that if $\Xset$ satisfies the manifold hypothesis, meaning informally that it has intrinsic dimension $m < d$ (see~\ref{asmp:domain_manifold} in Section~\ref{sec:graph_sobolev_classes} for a formal statement), both~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$ \citep{bickel2007,ingster2000}.
+Encouragingly, the smoothing spline estimator $\wt{f}$ and test $\wt{\phi}$ achieve these faster rates, with one caveat: in order for smoothing splines to properly leverage the manifold hypothesis, the domain $\Xset$ must be known.
 
-	\textcolor{red}{(TODO)}: Include references.
-
-	\item Finally, it is well known that if $\Xset$ satisfies the manifold hypothesis, meaning informally that it has intrinsic dimension $m < d$ (see~\ref{asmp:domain_manifold} in Section~\ref{sec:graph_sobolev_classes} for a formal statement), both~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
-Encouragingly, the smoothing spline estimator $\wt{f}$ and test $\wt{\phi}$ achieve these faster rates, with one caveat: in order for smoothing splines to properly leverage the manifold hypothesis, the domain $\Xset$ must be known.
-\end{itemize}
-
-\paragraph{(2.d) Consequences?} Intuitively, the connections between Laplacian smoothing and smoothing splines give us reason to hope that the strong statistical properties of the latter might be shared by the former.
+\paragraph{Consequences?} Intuitively, the connections between Laplacian smoothing and smoothing splines give us reason to hope that the strong statistical properties of the latter might be shared by the former.
 Unfortunately, it is not so easy to mathematically tie $\wh{f}$ to $\wt{f}$: for example, note that the convergence in~\eqref{eqn:seminorm_convergence} assumes $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:smoothing_spline} is either $H^1(\Xset)$ or $C^1(\Xset)$.
@@
 
-\textbf{(3)}  For $M > 0$, define the (first-order) graph Sobolev ball to be
+For $M > 0$, define the (first-order) graph Sobolev ball to be
 \begin{equation}
@@
 
-To understand why graph Sobolev balls play a critical role in the analysis of Laplacian smoothing, let us return for a moment to the continuous-time setting.
+To understand the role graph Sobolev balls play in our analysis of Laplacian smoothing, let us return for a moment to the continuous-time setting.
 Upper bounds on the error of smoothing splines rest on two crucial facts, both of which involve the continuum Laplace operator $\Delta_P$.
-The first is the \textit{a priori} assumption that the Sobolev semi-norm $|f|_{H^1(\Xset)} = \langle f, \Delta_P f \rangle_{\Leb^2(P)} \leq M^2$.\footnote{The equality holding when $P$ is uniform and  only for functions $f$ which satisfy boundary conditions.} The second is an upper bound on the metric entropy of $H^1(\Xset)$, characterized by the eigenvalue decay $\lambda_k(\Delta_P) \asymp k^{2/d}$ (i.e.
+The first is the \textit{a priori} assumption that the Sobolev semi-norm $|f|_{H^1(\Xset)}^2 = \langle f, \Delta_P f \rangle_{\Leb^2(P)} \leq M^2$.\footnote{The equality holding when $P$ is uniform and  only for functions $f \in C^2(\Xset)$ which satisfy boundary conditions.} The second is an upper bound on the metric entropy of $H^1(\Xset,M)$, characterized by the eigenvalue decay $\lambda_k(\Delta_P) \asymp k^{2/d}$ (i.e.
 Weyl's Law).
-In this section we will show that both these statements have analogues in the graph setting:
+In this section we will show that both these statements have analogues in the graph setting.
 \begin{itemize}
-	\item for functions $f$ belonging to $H^1(\Xset,M)$, the evaluations $f \in (f(X_1),\ldots,f(X_n)) \in \Reals^n$ are upper bounded,
+	\item For functions $f$ belonging to $H^1(\Xset,M)$, the evaluations $f \in (f(X_1),\ldots,f(X_n)) \in \Reals^n$ are upper bounded,
 	\begin{equation}
@@
 	\end{equation}
-	\item for a range $1 \leq k \leq k_{\max}(n)$, the eigenvalues $\lambda_k(G_{n,r})$ satisfy
+	\item For a range $1 \leq k \leq k_{\max}(n)$, the eigenvalues $\lambda_k(G_{n,r})$ satisfy
 	\begin{equation}
@@
 \end{itemize}
-Both statements will hold with high probability, and when we assume $\Xset$ is a manfiold of dimension $m$, they will hold with $d$ replaced by $m$.
+Both statements will hold with high probability, and when we assume $\Xset$ is a manifold of intrinsic dimension $m$, they will hold with $d$ replaced by $m$.
 These results elucidate why Laplacian smoothing ``works'' as well as its continuous-time counterpart ---in the sense of being statistically rate-optimal over (continuum) Sobolev classes.
 
-\paragraph{(3.a) Graph Sobolev semi-norm.}
+\paragraph{Graph Sobolev semi-norm.}
 We begin by upper bounding $f_0^T \Lap_{n,r} f_0$ under the assumption $f_0 \in H^1(\Xset,M)$. We will assume the kernel $K$ satisfies condition~\ref{asmp:kernel}.
@@
 	\label{asmp:kernel}
-	$K:[0,\infty) \to [0,\infty)$ is supported on $[0,1]$, and its restriction to $[0,1]$ is Lipschitz. Additionally, it is normalized so that
+	$K:[0,\infty) \to [0,\infty)$ is a non-increasing function supported on $[0,1]$, its restriction to $[0,1]$ is Lipschitz, and $K(1) > 0$. Additionally, it is normalized so that
 	\begin{equation*}
@@
 	\end{equation*}
+	Let $\sigma_K := \frac{1}{d} \int_{\Rd} \|x\|^2 K(\|x\|) \,dx$.
 \end{enumerate}
@@
 	\label{lem:graph_sobolev_seminorm}
-	Assume~\ref{asmp:kernel} and additionally that $p(x) \leq p_{\max}$ for all $x \in \Xset$.
-Then, for any $f \in H^1(\Xset)$, there exists a constant $C_1 > 0$ which depends only on $\Xset$, $d$, $p_{\max}$ and $K$ such that
+	Assume~\ref{asmp:kernel} and additionally that $p(x) \leq p_{\max}$ for all $x \in \Xset$. Then, for any $f \in H^1(\Xset)$, 
 	\begin{equation}
 	\label{eqn:graph_sobolev_seminorm}
-	f^T \Lap_{n,r} f \leq \frac{C_1}{\delta} n^2 r^{d + 2} |f|_{H^1(X)}^2
+	f^T \Lap_{n,r} f \leq \frac{p_{\max}^2 \sigma_K}{\delta} n^2 r^{d + 2} |f|_{H^1(X)}^2
 	\end{equation}
@@
 	The inequality~\eqref{eqn:graph_sobolev_seminorm} resembles~\eqref{eqn:seminorm_convergence}.
-However, as we pointed out previously to derive~\eqref{eqn:seminorm_convergence} one needs $f \in C^{2}(\Xset)$ as opposed to $f \in H^1(\Xset)$, and imposing this extra degree of regularity is unsuitable for our purposes.
-We use a more careful analysis to prove Lemma~\eqref{lem:graph_sobolev_seminorm}, but note well that~\eqref{eqn:graph_sobolev_seminorm} is only an upper bound rather than a statement about consistency.
+However, as we pointed out previously, to derive~\eqref{eqn:seminorm_convergence} one needs $f \in C^{2}(\Xset)$ as opposed to $f \in H^1(\Xset)$, and imposing this extra degree of regularity is unsuitable for our purposes.
+We use a more careful analysis to prove Lemma~\ref{lem:graph_sobolev_seminorm}, but note well that~\eqref{eqn:graph_sobolev_seminorm} is only an upper bound rather than a statement about consistency.
 This is the price we pay for insisting that $f$ have only one derivative; fortunately an upper bound will be sufficient for our purposes in Section~\ref{sec:minimax_optimal_laplacian_smoothing}.
@@
 
-\textbf{(3.b) Graph Laplacian eigenvalues.} The graph eigenvalues $\lambda_k(\Lap_{n,r}) \asymp n r^{d + 2}k^{2/d}$---when $\Xset$ is full-dimensional---or $\lambda_k(\Delta_P) \asymp nr^{m + 2}k^{2/m}$---when $\Xset$ has intrinsic dimension $1 < m < d$.
+\paragraph{Graph Laplacian eigenvalues.} Lemma~\ref{lem:neighborhood_eigenvalue} establishes upper and lower bounds on the graph Laplacian eigenvalues $\lambda_{k}(G_{n,r})$ in terms of the continuum Laplace eigenvalues $\lambda_k(\Delta_P)$.
+These bounds hold with high probability under appropriate regularity conditions on the domain $\Xset$, and for a range of radii $r$.
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{1}
+	\item
+	\label{asmp:domain}
+	The domain $\Xset$ is an open, connected set with Lipschitz boundary.
+\end{enumerate}
+\begin{enumerate}[label=(R\arabic*)]
+	\item
+	\label{asmp:radius1}
+	The radius $r$ satisfies $(\log n/n)^{1/d} \leq r \leq c_1$.
+\end{enumerate}
+\begin{lemma}
+	\label{lem:neighborhood_eigenvalue}
+	Suppose the density $p$, the domain $\Xset$, the kernel $K$ and the radius $r$ satisfy~\ref{asmp:bounded_lipschitz_density}, \ref{asmp:domain}, \ref{asmp:kernel},and~\ref{asmp:radius1} respectively.
+The following statement holds with probability at least $1 - A_1n\exp(-a_1nr^{d})$: for any $\ell \in \mathbb{N}$ such that
+	\begin{equation}
+	\label{eqn:neighborhood_eigenvalue_1}
+	1 - A_1 r \sqrt{\lambda_{\ell}(\Delta_P)} \geq \frac{1}{2}
+	\end{equation}
+	we have that
+	\begin{equation}
+	\label{eqn:eigenvalue_bound}
+	a_1 \lambda_k(G_{n,r}) \leq nr^{d+2} \lambda_k(\Delta_P) \leq A_1 \lambda_k(G_{n,r}),~~\textrm{for all $1 \leq k \leq \ell$}
+	\end{equation}
+\end{lemma}
+
+The scaling~\eqref{eqn:neighborhood_graph_eigenvalue_vague} then follows by Weyl's Law (see Lemma 28 of~\citep{dunlop2020} for a proof that~\ref{asmp:bounded_lipschitz_density} and~\ref{asmp:domain} are sufficient for Weyl's Law to hold.).
+\begin{corollary}
+	\label{cor:neighborhood_eigenvalue}
+	Suppose the density $p$ and the domain $\Xset$ satisfy~\ref{asmp:bounded_lipschitz_density} and~\ref{asmp:domain}; then there exist constants $a_2$ and $A_2$ such that
+	\begin{equation}
+	\label{eqn:weyls_law}
+	a_2k^{2/d} \leq \lambda_k(\Delta_P) \leq A_2k^{2/d}~~\textrm{for all $k \in \mathbb{N}, k > 1$}.
+	\end{equation}
+	Therefore under the conditions of Lemma~\ref{lem:neighborhood_eigenvalue}, we have that
+	\begin{equation}
+	\label{eqn:neighborhood_eigenvalue_2}
+	a_3 \cdot \min\Bigl\{nr^{d + 2}k^{2/d},nr^d\Bigr\} \leq \lambda_k(G_{n,r}) \leq A_3 \cdot \min\Bigl\{nr^{d + 2}k^{2/d}, nr^d\Bigr\}~~\textrm{for all $2 \leq k \leq n$}
+	\end{equation}
+	with probability at least $1 - A_1n\exp(-a_1nr^d)$.
+\end{corollary}
+
+We see that the inequalities in~\eqref{eqn:neighborhood_eigenvalue_2} match the desired scaling~\eqref{eqn:neighborhood_graph_eigenvalue_vague} for all $2 \leq k \leq (1/r)^d$.
+Indeed, if we could choose $r = n^{-1/d}$, we would get the desired scaling for all $2 \leq k \leq n$, but unfortunately choosing $r$ to be so small would violate~\ref{asmp:radius1}.
+Nevertheless, by choosing $r$ to be relatively small while still being in accordance with~\ref{asmp:radius1}, we can guarantee that only some of the largest eigenvalues $\lambda_k(G_{n,r})$ will fail to satisfy~\eqref{eqn:neighborhood_graph_eigenvalue_vague}.
+Fortunately, the behavior of these large eigenvalues will not meaningfully contribute to the error of $\wh{f}_{\LS}$ or $T_{\LS}$, as we shall see.
+
+\paragraph{Manifold hypothesis.}
+
+As mentioned in the introduction, an attractive quality of the neighborhood graph $G_{n,r}$ is that it captures the geometry of $P$ and $\Xset$.
+We now review two manifestations of this phenomenon, from \cite{trillos2019} and \cite{calder2019}: both the graph Sobolev semi-norm and graph Laplacian eigenvalues automatically adapt to the intrinsic dimension of $\Xset$.
+Formally, we make the following assumption on $\Xset$ (in order to directly use the results of~\cite{trillos2019} and~\cite{calder2019}).
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{2}
+	\item 
+	\label{asmp:domain_manifold}
+	The measure $P$ is supported on a compact, connected, smooth submanifold $\Xset$ of $\Reals^d$; the domain $\Xset$ is of fixed dimension $1 \leq m \leq d$ and without boundary.
+Moreover, $P$ has a density $p$ with respect to the volume form of $\Xset$ which is bounded away from $0$ and $\infty$,
+	\begin{equation*}
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty
+	\end{equation*}
+	for all $x \in \Xset$, and the density $p$ is Lipschitz with Lipschitz constant $L_p$.
+\end{enumerate}
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{1}
+	\item 
+	\label{asmp:radius2}
+	The radius $r$ satisfies $(\log n/n)^{1/m} \leq r \leq c_2$.
+\end{enumerate}
+\begin{proposition}[Lemma 5 of~\cite{trillos2019}]
+	\label{prop:manifold_graph_sobolev_seminorm}
+	Assume~\ref{asmp:kernel},~\ref{asmp:domain_manifold} and~\ref{asmp:radius2}. Then for any $f \in H^1(\Xset)$,
+	\begin{equation*}
+	f^T \Lap_{n,r} f \leq \frac{(1 + L_pr)^2}{\delta}n^{2}r^{m + 2} \sigma_K |f|_{H^1(\Xset)}^2
+	\end{equation*}
+	with probability at least $1 - \delta$.
+\end{proposition}
+
+\begin{proposition}[Restatement of Theorem 4 of~\cite{trillos2019}]
+	\label{prop:manifold_neighborhood_eigenvalue}
+	Assume~\ref{asmp:kernel}, ~\ref{asmp:domain_manifold}, and~\ref{asmp:radius2}. Then there exist constants $A_4, A_5$ and $a_4,a_5$ which do not depend on $n$ or $r$ such that 
+	\begin{equation}
+	\label{eqn:manifold_neighborhood_eigenvalue}
+	a_5 \cdot \min\Bigl\{nr^{m + 2}k^{2/m},nr^d\Bigr\} \leq \lambda_k(G_{n,r}) \leq A_5 \cdot \min\Bigl\{nr^{m + 2}k^{2/m}, nr^m\Bigr\}~~\textrm{for all $2 \leq k \leq n$}	
+	\end{equation}
+	with probability at least $1 - A_4n\exp(-a_4nr^m)$.
+\end{proposition}
+Note that~\ref{asmp:domain_manifold} assumes $\Xset$ is without boundary, and so Propositions~\ref{prop:manifold_graph_sobolev_seminorm} and~\ref{prop:manifold_neighborhood_eigenvalue} do not imply Lemmas~\ref{lem:graph_sobolev_seminorm} or~\ref{lem:neighborhood_eigenvalue}.
 
@@
 
-\textbf{(4)} We now leverage the results of Section~\ref{sec:graph_sobolev_classes} to formalize the main conclusions of this work: that Laplacian smoothing methods are minimax rate-optimal.
+We now leverage the results of Section~\ref{sec:graph_sobolev_classes} to formalize the main conclusions of this work: that Laplacian smoothing methods are minimax rate-optimal.
 
-\textbf{(4.a) Estimation error of Laplacian smoothing.} Using \textbf{(3.a)} and \textbf{(3.b)}, we establish upper bounds on the estimation error of Laplacian smoothing.
+For all results in this section, we will assume that the density~$p$ satisfies~\ref{asmp:bounded_lipschitz_density} and the kernel $K$ satisfies~\ref{asmp:kernel}.
+Additionally, within our theorem statements we will use $c$ and $C$ to denote constants which possibly differ from theorem to theorem, and which may depend on $K$, $d$, $p$, and $\Xset$ but do not depend on $f_0$, the radius of the Sobolev ball $M$, or the sample size $n$.
 
-\textbf{(4.b) Testing error of Laplacian smoothing.} Using \textbf{(3.a)} and \textbf{(3.b)}, we establish upper bounds on the testing error of Laplacian smoothing.
+\paragraph{Estimation error of Laplacian smoothing.} 
+Suppose the domain $\Xset$ satisfies~\ref{asmp:domain}.
+Under appropriate conditions, the Laplacian smoothing estimator $\wh{f}_{\LS}$ achieves, up to constants, the minimax risk $n^{-2/(2 + d)}$ over all $f_0 \in H^1(\Xset,M)$.
+To prove this, we will invoke Lemma~\ref{lem:graph_sobolev_seminorm} and Corollary~\ref{cor:neighborhood_eigenvalue} to suitably upper bound the squared-bias and variance of $\wt{f}_{\LS}$, respectively.
+In order to do so, we must restrict the range of the neighborhood graph radius $r$.
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{2}
+	\item 
+	\label{asmp:ls_kernel_radius_estimation}
+	The neighborhood graph radius $r$ satisfies
+	\begin{equation*}
+	\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq r \leq n^{-3/(4 + 2d)} M^{(d - 4)/(4 + 2d)}.
+	\end{equation*}
+\end{enumerate}
+\begin{theorem}
+	\label{thm:laplacian_smoothing_estimation1}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d < 4$ and $f_0 \in H^1(\Xset,M)$.
+Suppose that that the neighborhood graph $G_{n,r}$ is computed with a radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_estimation},  and the Laplacian smoothing estimator $\wh{f}_{\LS}$ with $\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$.
+It holds that
+	\begin{equation*}
+	\frac{1}{n}\Bigl\|\wh{f}_{\LS} - f_0\Bigr\|_2^2 \leq \frac{C}{\delta} M^{d/(2 + d)} n^{-2/(2 + d)}
+	\end{equation*}
+	with probability at least $1 - \delta -  A_1n\exp(-a_1nr^d) - \exp(-c M^{d/(2d + 4)} n^{d/(2+d)})$.
+\end{theorem}
+To summarize: when $d = 1,2$ or $3$, the Laplacian smoothing estimator $\wh{f}_{\LS}$ has in-sample mean squared error with a constant factor of the minimax risk, except on a set of small probability.
+Some remarks:
+\begin{itemize}
+	\item When $d = 4$, our analysis results in an upper bound on the mean squared error within a $\log n$ factor of the minimax risk, but when $d \geq 5$ our analysis yields upper bounds that are much worse than the minimax rate.
+This mirrors the conclusions of~\cite{sadhanala16}, who investigate estimation rates of Laplacian smoothing over the $d$-dimensional grid graph.
+In our setting, when $d = 4$ we would set $r \asymp (\log n/n)^{1/4}$---slightly outside the scaling~\ref{asmp:ls_kernel_radius_estimation}---to get the near-optimal rate.
+	\item The range of radii $r$ imposed by~\ref{asmp:ls_kernel_radius_estimation} is  compatible with practice, where by far the most common choice of radius is the connectivity threshold $r \asymp (\log(n)/n)^{1/d}$, chosen to make the graph $G_{n,r}$ as sparse as possible while still being connected.
+	\item The tuning parameter $\rho$ is chosen to balance the (conditional on ${\bf X}$) squared-bias and variance, i.e. chosen as large as possible while still with high probability satisfying
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_bias_variance}
+	\frac{\rho}{n} f_0^T \Lap_{{n,r}} f_0 \lesssim \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k}(G_{n,r}) + 1\bigr)^2}.
+	\end{equation}
+	To balance bias and variance as in~\eqref{eqn:laplacian_smoothing_bias_variance} we require control on all the graph eigenvalues $\lambda_1(G_{n,r}),\ldots,\lambda_n(G_{n,r})$.
+As mentioned in our commentary after Lemma~\ref{lem:neighborhood_eigenvalue}, when $r$ is small enough the large eigenvalues $\lambda_k(G_{n,r})$ contribute negligibly to the variance term.
+The upper bound in~\ref{asmp:radius1} is the largest value of $r$ such that this statement is still true.
+	\item The Laplacian smoothing estimator $\wh{f}_{\LS}$ is defined only at the design points $X_1,\ldots,X_n$ and we therefore measure error using the in-sample mean squared error $\frac{1}{n}\norm{\cdot}_2^2$.
+Because we work in the random design framework, the in-sample error is a random variable, and our bound is thus a bound in high probability.
+That being said, it is possible to smoothly extend $\wh{f}_{\LS}$ to be defined on all of $\Xset$, or indeed all of $\Reals^d$---for instance, using the \emph{Nystrom extension}---and evaluate the error of the extension in $\Leb^2(\Xset)$ norm.
+Assuming that the extension of our estimators and the regression function $f_0$ are suitably smooth, tools from empirical process theory will guarantee that the $\Leb^2(\Xset)$ error is not too much greater than the in-sample error, but we do not further pursue the details here.
+\end{itemize}
 
-\quad \textbf{(4.b.i) Low-smoothness regime.} Unlike in the estimation setting, the nonparametric goodness-of-fit testing problem itself exhibits an elbow when $d = 4$ and $f_0 \in \Leb^4(\Xset)$.
-The naive statistic $\frac{1}{n}\|Y\|_2^2$ is now optimal.
-By tuning Laplacian smoothing to the interpolation limit---i.e.
-setting $\rho = 0$---we recover this naive statistic, and therefore Laplacian smoothing is optimal in this sense (although really, no smoothing is being done.)
+\paragraph{Testing error of Laplacian smoothing.}
+Let us define a test using the statistic $T_{\LS}$. For $b \geq 1$, define the threshold $\wh{t}_b$ to be
+\begin{equation*}
+\wh{t}_{b} := \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k(G_{n,r}) + 1\bigr)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k(G_{n,r}) + 1\bigr)^4}},
+\end{equation*}
+and the Laplacian smoothing test $\wh{\phi}_{\LS} := \1\bigl\{T_{\LS} \leq \wh{t}_b\bigr\}$.
+(Clearly the test $\wh{\phi}_{\LS}$ depends on $b$, but we suppress this notationally.) The number $b \geq 1$ is selected by the user, with the choice made based on the tolerated level of Type I and Type II error.
+In fact, the threshold $\wh{t}_b$ is precisely the right choice to control the Type I error of $\wh{\phi}_{\LS}$; as we show in Lemma~\ref{lem:ls_fixed_graph_testing} in the supplementary material,
+\begin{equation*}
+\Ebb_0\Bigl[\wh{\phi}_{\LS}\Bigr] \leq \frac{1}{b^2}.
+\end{equation*}
 
-\textbf{(4.c) Comparing with smoothing splines.} While we motivated graph Laplacians by suggesting they might replicate the statistical properties of smoothing splines, we have in fact managed to show something stronger.
-In particular:
+In Theorem~\ref{thm:laplacian_smoothing_testing}, we show this test also has small Type II error, uniformly over all $f_0$ separated from $0$ by at least the critical radius $\epsilon\bigl(H^1(\Xset,M)\bigr)$ given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
+For this to hold, we will require a tight range of scalings for $r$.
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{3}
+	\item 
+	\label{asmp:ls_kernel_radius_testing}
+	The neighborhood graph radius $r$ satisfies
+	\begin{equation*}
+	\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq r \leq M^{(16 - d)/(8 + 2d)}n^{(d - 20)/(32 + 8d)}.
+	\end{equation*}
+\end{enumerate}
 
-\quad \textbf{(4.c.i)} When $d \geq 2$, there is no Sobolev embedding of $H^1(\Xset) \subseteq C^s(\Xset)$ for any $s > 0$. Therefore the smoothing spline estimator is ill-posed.
+\begin{theorem}
+	\label{thm:laplacian_smoothing_testing}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d < 4$ and $f_0 \in H^1(\Xset,M)$ for $M \leq n^{(4 - d)/(4 + d)}$.
+Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}, and the Laplacian smoothing test $\wh{\phi}_{\LS}$ with $\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$, and $b \geq 1$.
+For all $f_0$ such that
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_testing}
+	\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2d/(4 + d)} n^{-4/(4 + d)}
+	\end{equation} 
+	the Type II error is upper bounded:
+	\begin{equation*}
+	\Ebb_{f_0}\Bigl[1 - \wh{\phi}_{\LS} \Bigr] \leq C\biggl(\frac{1}{b}\Bigl[1 + n^{-d(4 + d)}M^{-2d/(4 + d)}\Bigr] + A_1n\exp\bigl(-a_1nr^d\bigr)\biggr).
+	\end{equation*}
+\end{theorem}
 
-\quad \textbf{(4.c.ii)} Additionally, the metric entropy of $C^1(\Xset)$ is too large to obtain optimal rates for \textcolor{red}{Holder-smoothing splines} when $d \geq 2$.
-Thus, we see a gap emerge between the optimality properties of the discrete-time (Laplacian smoothing) and continuous-time (smoothing splines) estimators, in favor of the latter.
+A few remarks:
+\begin{itemize}
+	\item As mentioned in Section~\ref{sec:minimax_optimal_regression_sobolev_spaces}, when $d \geq 4$ the Sobolev balls $H^1(\Xset,M)$ include quite irregular functions $f \not\in \Leb^4(\Xset)$.
+Proving tight lower bounds over such classes is non-trivial, and to the best of our knowledge such an analysis remains outstanding.
+On the other hand, if we explicitly assume that $f \in \Leb^4(\Xset,M)$, then \cite{guerre02} show that the testing problem is characterized by the dimension-free lower bound $\epsilon^{\star}(\Leb^4(\Xset,M)) \gtrsim n^{-1/4}$.
+Moreover, by training $\wh{f}_{\LS}$ to the interpolation limit, i.e.
+setting $\rho = 0$, the test $\wh{\phi}_{\LS}$ achieves (up to constants) this lower bound.
+That is,
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_testing_low_smoothness}
+	E_{f_0}\Bigl[1 - \wh{\phi}_{\LS}\Bigr] \leq \frac{C(1 + M^4)}{b^2}
+	\end{equation}
+	for any $f_0 \in \Leb^4(\Xset,M)$ for which $\|f_0\|_{\Leb^2(\Xset)} \geq C bn^{-1/4}$. 
+	\item To compute the threshold $\wh{t}_b$, one must know each of the eigenvalues $\lambda_1(G_{n,r}),\ldots,\lambda_n(G_{n,r})$.
+Computing all $n$ of these eigenvalues is far more expensive than computing $T_{\LS}$.
+That being said, in practice we would not recommend using $\wh{t}_b$ anyway: the threshold is justified by asymptotic theory, but in practice $n$ is always finite, and using such thresholds can result in substantial loss of efficiency.
+We would instead make the standard recommendation to calibrate via permutation.
+\end{itemize}
+It is worth circling back to our comparison of Laplacian smoothing methods to smoothing splines.
+While we motivated Laplacian smoothing methods in part by suggesting they might replicate the statistical properties of smoothing splines, we have in fact managed to show something stronger.
+Theorems~\ref{thm:laplacian_smoothing_estimation1} and~\ref{thm:laplacian_smoothing_testing} imply that Laplacian smoothing methods are minimax optimal for both estimation and testing in dimensions $d = 1,2$ and $3$, and are nearly minimax-optimal for the estimation problem when $d = 4$.
+On the other hand, as covered in Section~\ref{sec:minimax_optimal_regression_sobolev_spaces} smoothing spline methods are optimal only when $d = 1$ or, for Lipschitz smoothing splines, $d = 2$.
+Thus, we see a somewhat surprising gap emerge between the optimality properties of the discrete-time (Laplacian smoothing) and continuous-time (smoothing spline) approaches, in favor of the former.
 
-\quad \textbf{(4.c.iii)} On the other hand, when $d \geq 4$, we no longer get the known optimal rates in the estimation problem.
-When $d = 4$, our rates are suboptimal by a factor of $\log n$.
-When $d > 5$ they are very suboptimal.
-We believe this is related to classical results regarding the entropy of low-smoothness Holder classes, i.e.
-the results used to justify \textbf{(4.c.ii)}.
-The interesting point is that the parameter spaces $H^1(\Xset)$, $C^1(\Xset)$ and $H^1(G_{n,r})$ become ``too large'' for different values of $d$.
+\paragraph{Manifold adaptivity.}
 
-\quad \textbf{(4.c.iv)} \textcolor{red}{(TODO)}: Translate these results into error in $\Leb^2(P)$ error, if/when possible.
+In Section~\ref{sec:graph_sobolev_classes}, we reviewed that when $\Xset$ is a submanifold of $\Rd$, the graph Laplacian automatically adapts to the geometry---and in particular, the intrinsic dimension $m$---of $\Xset$.
+This translates into improved upper bounds on the estimation and testing error of Laplacian smoothing; the upper bounds now depend on $m$ instead of $d$.
 
-\textbf{(4.d) Manifold adaptivity.} In Section~\ref{sec:graph_sobolev_classes}, we established that when $\Xset$ is an $m$-dimensional submanifold of $\Reals^d$, Laplacian smoothing approaches automatically ``feel'' the intrinsic dimension $m$ of $\Xset$.
-We now show that this translates into improved upper bounds on estimating and testing error.
+\begin{theorem}
+	\label{thm:laplacian_smoothing_manifold}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
+Suppose that $\Xset$ satisfies~\ref{asmp:domain_manifold} for $m < 4$, and that $f_0 \in H^1(\Xset,M)$.
+For appropriate choices of tuning parameters $r$, $\rho$, and $b \geq 1$, each of the following statements holds.
+	\begin{itemize}
+		\item The Laplacian smoothing estimator satisfies
+		\begin{equation*}
+		\frac{1}{n}\bigl\|\wh{f}_{\LS} - f_0\bigr\|_2^2 \leq \frac{C}{\delta}M^{m/(2 + m)}n^{-2/(2 + m)}
+		\end{equation*}
+		with probability at least $1 - \delta - A_4n\exp(-a_4nr^m) - \exp(-c M^{m/(2m + 4)} n^{m/(2+m)})$.
+		\item Suppose $M \leq n^{(4 - m)/(4 + m)}$. For all $f_0$ such that
+		\begin{equation*}
+		\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2m/(4 + m)} n^{-4/(4 + m)}
+		\end{equation*} 
+		the type II error of $\wh{\phi}$ is upper bounded:
+		\begin{equation*}
+		\Ebb_{f_0}\Bigl[1 - \wh{\phi}_{\LS} \Bigr] \leq C\biggl(\frac{1}{b}\Bigl[1 + n^{-m(4 + m)}M^{-2m/(4 + m)}\Bigr] + A_4n\exp\bigl(-a_4nr^m\bigr)\biggr).
+		\end{equation*}
+	\end{itemize}
+\end{theorem}
+In our supplementary material, we specify the right choices of $\rho$, as well as the permissible ranges of $r$.
+For now, we merely point out that they will depend only on the intrinsic dimension $m$, and not on any other properties of the manifold $\Xset$.
+In all other respects, $\wh{f}_{\LS}$ and $T_{\LS}$ are computed without any knowledge of $\Xset$.
+By contrast, the penalty in~\eqref{eqn:smoothing_spline} must be specially tailored to $\Xset$ in order for $\wt{f}_{\SM}$ and $T_{\SM}$ to achieve the minimax estimation and testing rates.
+This reveals another advantage of Laplacian smoothing as opposed to smoothing splines.
 
-\quad \textbf{(4.d.i)} This confirms the statement in \textbf{(4)} that Laplacian smoothing automatically adapts to the geometry of $\Xset$ and $p$.
+\section{Simulations}
+\label{sec:simulations}
 
-\quad \textbf{(4.d.ii)} By contrast, smoothing splines must be specially tailored to work in this setting.
+\textcolor{red}{(TODO)}
 
-\textbf{(4.e) Higher-order smoothness classes.} Laplacian smoothing can also be adapted to take advantage of additional assumed regularity on $f_0$, i.e.
+\section{Discussion}
+\label{sec:discussion}
+
+\textcolor{red}{(TODO)}
+
+A few discussion items:
+\begin{itemize}
+	\item \textbf{Higher-order smoothness classes.} Laplacian smoothing can also be adapted to take advantage of additional assumed regularity on $f_0$, i.e.
 $f_0 \in H^s(\mc{\Xset})$ for $s > 1$.
 In the very special case of $s = 2$ and $d = 2$, we can show that this adapted estimator achieves the sharper minimax rates over these classes, but the general story--for all combinations of $s$ and $d$---remains beyond our reach.
-
-\section{Simulations}
-
-\textbf{(5)} Empirically, we demonstrate that the risk for Laplacian smoothing is comparable to that of smoothing splines---in the full dimensional case with uniformly random design.
-When the design is sampled from a density concentrated near a manifold, Laplacian smoothing outperforms (out of the box) smoothing splines.
-
-\section{Discussion}
-
-\textcolor{red}{(TODO)}
+	\item \textbf{Possible extensions}. Methodologically: kNN graphs and different normalizations of the Laplacian. Theoretically: relaxing~\ref{asmp:bounded_lipschitz_density}. 
+\end{itemize}
 
@@
 
-Many of our results follow the same general two-part proof strategy.
-First, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to functionals of the graph $G$.
-We then analyze the behavior of these functionals with respect to the particular neighborhood graph $G_{n,r}$ and give high probability (upper or lower) bounds on these functionals.
+Our results all follow the same general two-part proof strategy, centered on conditioning.
+First, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to functionals of the graph $G$, and allow us to upper bound the error of $\wh{f}_{\LS}$ and $T_{\LS}$ conditional on $\mathbf{X} = x$.
+Second, we analyze the behavior of these functionals with respect to the particular neighborhood graph $G_{n,r}$ and give high probability (upper or lower) bounds.
 It is in this second step where we invoke our various assumptions on the distribution $P$ and regression function $f_0$.
@@
 
+Suppose we observe a fixed graph $G = \bigl([n],W\bigr)$ with Laplacian $\Lap_G$ and responses
+\begin{equation}
+\label{eqn:fixed_graph_regression_model}
+Y_i = f_{0,i} + \varepsilon_i
+\end{equation}
+where $f_0 = (f_{0,1},\ldots,f_{0,n}) \in \Reals^n$, and the noise variables $\varepsilon_i$ are independent $N(0,1)$. The Laplacian smoothing estimator of $f_0$ on $G$ is
+\begin{equation}
+\label{eqn:ls_G}
+\wh{f}_{\LS}(G) := \argmin_{f \in \Reals^n} \biggl\{ \sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_G^{s}  f \biggr\} = (\rho \Lap_G^s + \Id)^{-1}Y.
+\end{equation}
+The statistic
+\begin{equation}
+\label{eqn:ls_ts_G}
+T_{\LS}(G) := \frac{1}{n} \Bigl\|\wh{f}_{\LS}(G)\Bigr\|_2^2 
+\end{equation}
+is used to distinguish
+\begin{equation*}
+\mathbf{H}_0: f_{0} = (0,...,0) ~~\textrm{versus}~~ \mathbf{H}_a: f_{0} \neq (0,...,0).
+\end{equation*}
+We note that~\eqref{eqn:laplacian_smoothing} and \eqref{eqn:laplacian_smoothing_test} specialize~\eqref{eqn:ls_G} and \eqref{eqn:ls_ts_G} to the case where $G = G_{n,r}$ and $s = 1$.
+
+\subsection{Error bounds for linear smoothers}
+Let $S \in \Reals^{n \times n}$ be a square, symmetric matrix, and let $\check{f} = SY$ be a linear estimator of $f_0$.
+Clearly \eqref{eqn:ls_G} can be written in this form, and Lemma~\ref{lem:linear_smoother_fixed_graph_estimation} is therefore useful in analyzing it.
+Let $\lambdavec(S) = (\lambda_1(S),\ldots,\lambda_n(S)) \in \Reals^n$ denote the eigenvalues of $S$ and let $v_k(S)$ denote the corresponding \emph{unit-norm} eigenvectors, so that $S = \sum_{k = 1}^{n} \lambda_k(S) \cdot v_k(S) v_k(S)^T$.
+Denote $Z_k = v_k(S)^T \varepsilon$, and observe that $Z = (Z_1,\ldots,Z_n) \sim N(0,\Id)$.
+
+\begin{lemma}
+	\label{lem:linear_smoother_fixed_graph_estimation}
+	Let $\check{f} = SY$ for a square, symmetric matrix, $S \in \Reals^{n \times n}$ satisfying $\lambda_n(S) \leq 1$. Then
+	\begin{equation*}
+	\Pbb_{f_0}\biggl(\frac{1}{n}\bigl\|\check{f} - f_0\bigr\|_2^2 \geq \frac{10}{n} \bigl\|\lambdavec(S)\bigr\|_2^2 + \frac{2}{n}\bigl\|(S - I)f_0\bigr\|_2^2\biggr) \leq 1 - \exp\Bigl(-\bigl\|\lambdavec(S)\bigr\|_2^2\Bigr)
+	\end{equation*}
+\end{lemma}
+
+On the testing side, the functional $\norm{\check{f}}_2^2 = Y^T S^2 Y$ is a $U$-statistic of order $2$.
+The statistic $T_{\LS}(G)$ can be written in this form, and Lemma~\ref{lem:linear_smoother_fixed_graph_testing} is therefore useful in analyzing it.
+\begin{lemma}
+	\label{lem:linear_smoother_fixed_graph_testing}
+	Let $T = Y^T S^2 Y$ for a square, symmetric matrix $S \in \Reals^{n \times n}$. Define the threshold $t_b$ to be 
+	\begin{equation}
+	t_{b} := \norm{\lambdavec(S)}_2^2 + 2b \norm{\lambdavec(S)}_4^2
+	\end{equation}
+	Suppose the eigenvalues $0 \leq \lambda_{1}(S) \leq \lambda_{n}(S) \leq 1$. Then,
+	\begin{itemize}
+		\item \textbf{Type I error.}
+		\begin{equation}
+		\label{eqn:linear_smoother_fixed_graph_testing_typeI}
+		\Pbb_0\bigl(T > t_b\bigr) \leq \frac{1}{b^2}
+		\end{equation}
+		\item \textbf{Type II error.} Suppose further that
+		\begin{equation}
+		\label{eqn:linear_smoother_fixed_graph_testing_critical_radius}
+		f_0^T S^2 f_0 \geq 4b \norm{\lambdavec(S)}_4^2.
+		\end{equation}
+		Then
+		\begin{equation}
+		\label{eqn:linear_smoother_fixed_graph_testing_typeII}
+		\Pbb_{f_0}\bigl(T \leq t_b\bigr) \leq \frac{4}{b^2} + \frac{8}{b \norm{\lambdavec(S)}_4^{2}} 
+		\end{equation}
+	\end{itemize}
+\end{lemma}
+
+\paragraph{Proof of Lemma~\ref{lem:linear_smoother_fixed_graph_estimation}.}
+It holds that
+\begin{align*}
+\frac{1}{n}\bigl\|\check{f} - f_0\bigr\|_2^2 & \leq \frac{2}{n}\Bigl(\bigl\|\check{f} - \Ebb_{f_0}[\check{f}]\bigr\|_2^2 + \bigl\|\Ebb_{f_0}[\check{f}] - f_0\bigr\|_2^2\Bigr) \\ 
+& = \frac{2}{n}\Bigl(\bigl\|S\varepsilon\bigr\|_2^2 + \bigl\|(S - I)f_0\bigr\|_2^2\Bigr)
+\end{align*}
+Writing $\norm{S\varepsilon}_2^2 = \sum_{k = 1}^{n} \lambda_k(S)^2 Z_k^2$, the claim follows from the result of \cite{laurent00} on concentration of $\chi^2$-random variables, which for completeness we restate in Lemma~\ref{lem:chi_square_bound}.
+To be explicit, taking $t = \norm{\lambdavec(S)}_2^2$ in Lemma~\ref{lem:chi_square_bound} completes the proof of Lemma~\ref{lem:linear_smoother_fixed_graph_estimation}.
+
+\paragraph{Proof of Lemma~\ref{lem:linear_smoother_fixed_graph_testing}.}
+	We compute the mean and variance of $T$ as a function of $f_0$, then apply Chebyshev's inequality.
+
+	\textit{Mean.} Writing $Y = f_0 + \varepsilon$, we make use of the eigendecomposition of $S$ to obtain
+	\begin{equation}
+	\label{pf:linear_smoother_fixed_graph_testing1}
+	\begin{aligned}
+	T & = f_0^T S^2 f_0 + 2 f_0^T S^2 \varepsilon + \varepsilon^T S^2 \varepsilon \\
+	& = f_0^T S^2 f_0 + 2 f_0^T S^2 \varepsilon + \sum_{k = 1}^{n}  \bigl(\lambda_k(S)\bigr)^2 (\varepsilon^T v_k(S))^2 \\
+	& = f_0^T S^2 f_0 + 2 f_0^T S^2 \varepsilon + \sum_{k = 1}^{n}  \bigl(\lambda_k(S)\bigr)^2 Z_k^2,
+	\end{aligned}
+	\end{equation}
+	implying
+	\begin{equation}
+	\label{pf:linear_smoother_fixed_graph_testing_mean}
+	\Ebb_{f_0}[T] = f_0^T S^2 f_0 + \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^2.
+	\end{equation}
+
+	\textit{Variance.} Starting from~\eqref{pf:linear_smoother_fixed_graph_testing1} and recalling the basic fact $\Var(Z_k^2) = 2$, we derive
+	\begin{equation}
+	\label{pf:linear_smoother_fixed_graph_testing_var}
+	\Var_{f_0}[T] \leq 8 f_0^T S^4 f_0 + 4 \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^4 \leq 8 f_0^T S^2 f_0 + 4 \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^2
+	\end{equation}
+	where the second inequality follows since by assumption $\lambda_{n}(S) \leq 1$.
+
+	\textit{Bounding Type I and Type II error.} The bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeI} follows directly from Chebyshev's inequality, along with our above calculations on the mean and variance of $T$.
+
+	The bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeII} also follows from Chebyshev's inequality, as can be seen by the following manipulations,
+	\begin{equation*}
+	\begin{aligned}
+	\Pbb_{f_0}\bigl(T \leq t_b\bigr) & = \Pbb_{f_0}\bigl(T - \Ebb_{f_0}[T] \leq t_b - \Ebb_{f_0}[T]\bigr) \\
+	& \overset{(i)}{\leq} \Pbb_{f_0}\bigl(\abs{T - \Ebb_{f_0}[T]} \geq \abs{t_b - \Ebb_{f_0}[T]}\bigr) \\ 
+	& \overset{(ii)}{\leq} 4 \frac{\Var_{f_0}[T]}{(f_0^T S^2 f_0)^2} \\
+	& \overset{(iii)}{\leq} \frac{32}{f_0^T S^2 f_0} + \frac{4}{b^2} \\
+	& \overset{(iv)}{\leq} \frac{8}{b} \Biggl(\sum_{k = 1}^{n}\bigl(\lambda_k(S)\bigr)^4\Biggr)^{-1/2} + \frac{4}{b^2}
+	\end{aligned}
+	\end{equation*}
+	In the previous expression, $(i)$ and $(ii)$ follow since assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and equation~\eqref{pf:linear_smoother_fixed_graph_testing_mean} together imply $\Ebb_{f_0}(T) - \frac{1}{2}f_0^T S^2f_0 \geq t_b$, $(iii)$ follows from assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and the inequality~\eqref{pf:linear_smoother_fixed_graph_testing_var}, and $(iv)$ follows assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}.
+
+\subsection{Analysis of Laplacian Smoothing}
+In Lemma~\ref{lem:ls_fixed_graph_estimation}, we upper bound the mean squared error of $\wh{f}_{\LS}(G)$.
+\begin{lemma}
+	\label{lem:ls_fixed_graph_estimation}
+	For any $\rho,s > 0$,
+	\begin{equation}
+	\label{eqn:ls_fixed_graph_estimation_prob}
+	\frac{1}{n}\bigl\|\wh{f}_{\LS}(G) - f_0\bigr\|_2^2 \leq \frac{2\rho}{n} \bigl(f_0^T \Lap_{G}^s f_0\bigr) + \frac{10}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k}(G)^s + 1\bigr)^2}
+	\end{equation}
+	with probability at least $1 - \exp\Bigl(-\sum_{k = 1}^{n}\bigl(\rho \lambda_k(G)^s + 1\bigr)^{-2}\Bigr)$.
+\end{lemma}
+
+In Lemma~\ref{lem:ls_fixed_graph_testing}., we upper bound the testing error of a test based on $T_{\LS}(G)$.
+\begin{lemma}
+	\label{lem:ls_fixed_graph_testing}
+	Define the threshold $\wh{t}_b$ to be
+	\begin{equation*}
+	\wh{t}_b := \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k(G)^s + 1\bigr)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k(G)^s + 1\bigr)^4}}
+	\end{equation*}
+	Then each of the following holds for any $\rho,s > 0$ and any $b \geq 1$.
+	\begin{itemize}
+		\item \textbf{Type I error.}
+		\begin{equation}
+		\label{eqn:ls_fixed_graph_testing_typeI}
+		\Pbb_0\Bigl(T_{\LS}(G) > \wh{t}_b\Bigr) \leq \frac{1}{b^2}
+		\end{equation}
+		\item \textbf{Type II error.} If
+		\begin{equation}
+		\label{eqn:ls_fixed_graph_testing_critical_radius}
+		\frac{1}{n}\norm{f_0}_2^2 \geq \frac{2 \rho}{n} \bigl(f_0^T \Lap_G^s f_0\bigr) + \frac{4b}{n} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G)^s + 1)^4} \Biggr)^{1/2}
+		\end{equation}
+		then
+		\begin{equation}
+		\label{eqn:ls_fixed_graph_testing_typeII}
+		\Pbb_{f_0}\Bigl(T_{\LS}(G) \leq \wh{t}_b\Bigr) \leq \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G)^s + 1)^4} \Biggr)^{-1/2}
+		\end{equation}
+	\end{itemize}
+\end{lemma}
+
+\paragraph{Proof of Lemma~\ref{lem:ls_fixed_graph_estimation}.}
+Letting $\wh{S} = (\Id + \rho \Lap_G^s)^{-1}$, the estimator $\wh{f}_{\LS}(G) = \wh{S}Y$.
+The matrix $\wh{S} \in \Reals^{n \times n}$ is symmetric, and satisfies $\lambda_k(\wh{S}) \in (0,1)$ for all $k \in [n]$.
+The claim then follows from Lemma~\ref{lem:linear_smoother_fixed_graph_estimation} upon noting first that
+\begin{equation*}
+\bigl\|\lambdavec(\wh{S})\bigr\|_2^2 = \sum_{k = 1}^{n} \frac{1}{\bigl(1 + \rho \lambda_k(G)^s\bigr)^2}
+\end{equation*} 
+and second that
+\begin{equation*}
+\begin{aligned}
+\bigl\|(\wh{S} - I)f_0\bigr\|_2^2 & = f_0^T \Lap_G^{s/2} \Lap_G^{-s/2}\bigl(\wh{S} - \Id\bigr) \Lap_G^{-s/2} \Lap_G^{s/2} f_0 \\
+& \leq f_0^T \Lap_G^s f_0 \cdot \lambda_n\Bigl(\Lap_G^{-s/2}\bigl(\wh{S} - \Id\bigr)\Lap_G^{-s/2}\Bigr) \\
+& = f_0^T \Lap_G^s f_0 \cdot \max_{k \in [n]} \biggl\{\frac{1}{\lambda_k(G)^s} \Bigl(1 - \frac{1}{\rho\lambda_k(G)^s + 1}\Bigr) \biggr\} \\
+& = f_0^T \Lap_G^s f_0 \cdot \rho.
+\end{aligned}
+\end{equation*} 
+where we write $\Lap_G^{-1}$ for the pseudoinverse of $\Lap_G$.
+
+\paragraph{Proof of Lemma~\ref{lem:ls_fixed_graph_testing}.}
+Let $\wh{S} := (\rho \Lap^s + \Id)^{-1}$.
+The matrix $\wh{S} \in \Reals^{n \times n}$ is symmetric and positive semidefinite, and our test statistic $T_{\LS}(G) = \frac{1}{n}Y^T \wh{S}^2 Y$.
+The desired result thus follows from Lemma~\ref{lem:linear_smoother_fixed_graph_testing}.
+To see that the conditions of this Lemma are satisfied, we first note that since
+\begin{equation*}
+\lambda_k(\wh{S}) = \frac{1}{(\rho\lambda_k(G)^s + 1)}
+\end{equation*}
+and $\rho, \lambda_k(G) > 0$, it is evident that $\lambda_{n}(\wh{S}) \leq 1$.  Then, by assumption~\eqref{eqn:ls_fixed_graph_testing_critical_radius}
+\begin{equation*}
+f_0^{T} \wh{S}^2 f_0 = \norm{f_0}_2^2 - f_0^T(I - \wh{S}^2)f_0 \geq 2 \rho (f_0^T \Lap^s f_0) + f_0^T(I - \wh{S}^2)f_0 + 4b \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G)^s + 1)^4} \Biggr)^{-1/2},
+\end{equation*}
+and along with the following calculations,
+\begin{equation*}
+\begin{aligned}
+f_0^T \Bigl(\Id - \wh{S}^2\Bigr) f_0  & = f_0^T L^{s/2} L^{-s/2}\Bigl(\Id - \wh{S}^2\Bigr) L^{-s/2} L^{s/2} f_0 \\ 
+& \leq f_0^T L^{s} f_0 \cdot  \lambda_{\max}\biggl(L^{-s/2}\Bigl(\Id - \wh{S}^2\Bigr) L^{-s/2}\biggr) \\ 
+& \overset{(i)}{=}  f_0^T L^{s} f_0 \cdot \max_{k} \biggl\{ \frac{1}{\lambda_k(G)^s} \Bigl(1 - \frac{1}{(\rho \lambda_k(G)^s + 1)^2}\Bigr) \biggr\} \\
+& \overset{(ii)}{\leq} f_0^T L^{s} f_0 \cdot 2\rho,
+\end{aligned}
+\end{equation*}
+we have that
+\begin{equation*}
+f_0^{T} \wh{S}^2 f_0 \geq 2b \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G)^s + 1)^4} \biggr)^{-1/2}.
+\end{equation*} 
+In other words condition~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} in Lemma~\ref{lem:linear_smoother_fixed_graph_testing} is met, and applying that Lemma completes the proof.
+
+(In the previous derivation: in $(i)$ the maximum is over all indices $k$ such that the eigenvalue $\lambda_k(G)$ is strictly positive; and $(ii)$ follows from the basic algebraic identity $1 - 1/(1 + \rho x)^2 \leq 2 \rho x$ for any $x, \rho > 0$.
+
 \section{Neighborhood graph Sobolev semi-norm}
 \label{sec:graph_sobolev_seminorm}
+In this section, we prove Lemma~\ref{lem:graph_sobolev_seminorm}. To do so, we will show that for any $f \in H^1(\Xset)$,
+\begin{equation*}
+\Ebb\Bigl[f^T \Lap_{n,r} f\Bigr] \leq p_{\max}^2 \sigma_K n^2 r^{d + 2} |f|_{H^1(\Xset)}^2
+\end{equation*}
+whence the claim follows immediately by Markov's inequality (recall that $\Lap_{n,r}$ is positive semi-definite, and therefore $f^T \Lap_{n,r} f$ is a non-negative random variable).
+
+Since
+\begin{equation*}
+f^T \Lap_{n,r} f = \frac{1}{2}\sum_{i, j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 \mathbf{W}_{ij},
+\end{equation*}
+it follows that
+\begin{equation}
+\label{pf:first_order_graph_sobolev_seminorm_1}
+\Ebb\Bigl[f^T \Lap_{n,r} f\Bigr] = \frac{n(n - 1)}{2} \Ebb\biggl[\Bigl(f(X') - f(X)\Bigr)^2 K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr]
+\end{equation}
+where $X$ and $X'$ are random variables independently drawn from $P$.
+
+For the remainder of this proof, we will assume that $f \in C^{\infty}(\Xset)$, which we may do without loss of generality because $C^{\infty}(\Xset)$ is dense in $H^1(\Xset)$ and the expectation on the right hand side of~\eqref{pf:first_order_graph_sobolev_seminorm_1} is continuous in $\Leb^2(\Xset)$.
+Obviously
+\begin{equation}
+\Ebb\biggl[\bigl(f(X') - f(X)\bigr)^2K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr] \leq p_{\max}^2 \int_{\Xset} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K_r(x',x) \,dx' \,dx.
+\label{pf:first_order_graph_sobolev_seminorm_2}
+\end{equation}
+and it remains now to bound the integral. Taking a first-order Taylor expansion of $f$, we get
+\begin{align}
+\int_{\Xset} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K_r(x',x) \,dx' \,dx & = \int_{\Xset} \int_{\Xset} \biggl[\int_{0}^{1} \nabla f\bigl(x + t(x' - x)\bigr)^T (x' - x)\,dt\biggr]^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx' \,dx \nonumber \\
+& \overset{(i)}{\leq} \int_{\Xset} \int_{\Xset} \int_{0}^{1} \biggl[\nabla f\bigl(x + t(x' - x)\bigr)^T (x' - x)\biggr]^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dt\,dx' \,dx \nonumber \\
+& \overset{(ii)}{=} r^{d + 2} \int_{\Xset} \int_{\mc{Z}_r(x)} \int_{0}^{1} \biggl[\nabla f\bigl(x + trz\bigr)^T z\biggr]^2 K\bigl(\|z\|\bigr) \,dt\,dz \,dx \\
+&  \overset{(iii)}{\leq} r^{d + 2} \int_{\Xset} \int_{\Reals^d} \int_{0}^{1} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) \,dt \,dz \,d\wt{x} \nonumber \\
+& = r^{d + 2} \int_{\Xset} \int_{\Reals^d} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) \,dz \,d\wt{x} \label{pf:first_order_graph_sobolev_seminorm_3}
+\end{align}
+where $(i)$ follows by Jensen's inequality, $(ii)$ follows by substituting $z = (x' - x)/r$ with $\mc{Z}_r(x) := \{z: zr + x \in \Xset\}$ and $(iii)$ by substituting $\wt{x} = x + trz$ and noting that
+\begin{equation*}
+x \in \Xset ~~\textrm{and}~~ z \in \mc{Z}_r(x) \Longrightarrow x + trz \in \Xset ~~\textrm{and}~~ z \in \Reals^d.
+\end{equation*}
+Now, writing $\bigl(\nabla f(\wt{x}) ^T z\bigr)^2 = \bigl(\sum_{i = 1}^{d} z_{i} f^{(e_i)}(x) \bigr)^2$, expanding the square and integrating, we have that for any $\wt{x} \in \Xset$,
+\begin{align*}
+\int_{\Reals^d} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) \,dz & = \sum_{i,j = 1}^{d} f^{(e_i)}(\wt{x}) f^{(e_j)}(\wt{x}) \int_{\Rd} z_iz_jK(\|z\|) \,dz \\
+& = \sum_{i = 1}^{d} \bigl(f^{(e_i)}(\wt{x})\bigr)^2 \int_{\Rd} z_i^2 K\bigl(\|z\|\bigr) \,dz \\
+& = \sigma_K \|\nabla f(\wt{x})\|^2
+\end{align*}
+and by plugging back into~\eqref{pf:first_order_graph_sobolev_seminorm_3}, we obtain the claimed result.
 
@@
 \label{sec:graph_eigenvalues}
+In this section, we prove our claimed results regarding the eigenvalues $\lambda_k(G_{n,r})$.
+We begin with Corollary~\ref{cor:neighborhood_eigenvalue}, since its proof is straightforward, and then turn to the proof of Lemma~\ref{lem:neighborhood_eigenvalue}, which is more involved.
+
+\subsection{Proof of Corollary~\ref{cor:neighborhood_eigenvalue}}
+Put
+\begin{equation*}
+\ell_{\star} = \floor{\frac{\bigl(1/2 - A_1(\theta + \wt{\delta})\bigr)^d}{C_2^d A_1^d r^d}};
+\end{equation*}
+and note that for $\ell_{\star}$ the condition~\eqref{eqn:neighborhood_eigenvalue_1} is met, i.e.
+\begin{equation*}
+1 - A_1\Biggl(r \sqrt{\lambda_{\ell}(\Delta_P)} + (\theta + \wt{\delta})\Biggr)  \geq \frac{1}{2}.
+\end{equation*}
+We may therefore apply the conclusions of Lemma~\ref{lem:neighborhood_eigenvalue}, which along with~\eqref{eqn:weyls_law} imply that
+\begin{equation*}
+\frac{c_2}{A_1} nr^{d + 2} k^{2/d} \leq \lambda_k(G_{n,r}) \leq \frac{C_2}{a_1} nr^{d + 2} k^{2/d}~~\textrm{for all $2 \leq k \leq \ell_{\star}$}
+\end{equation*}
+On the other hand, if $k > \ell^{\star}$,...
+
+\textcolor{red}{(TODO)}: Finish off the proof.
+
+\subsection{Proof of Lemma~\ref{lem:neighborhood_eigenvalue}}
+
+In this section we prove Lemma~\ref{lem:neighborhood_eigenvalue}, closely following the lead of~\citep{burago2014,trillos2019,calder2019}.
+As in these works, we relate $\lambda_k(\Delta_P)$ and $\lambda_k(G_{n,r})$ by means of the Sobolev semi-norms (Dirichlet energies)
+\begin{equation*}
+b_r(u) := \frac{1}{n^2 r^{d+ 2}}u^T \Lap_{n,r} u~~\textrm{for any $u \in \Leb^2(P_n)$}
+\end{equation*}
+and
+\begin{equation*}
+D_2(f) :=
+\begin{cases*}
+\int_{\Xset} \|\nabla f(x)\|^2 p^2(x) \,dx~~ &\textrm{if $f \in H^1(\Xset)$} \\
+\infty~~ & \textrm{otherwise,}
+\end{cases*}
+\end{equation*}
+Let us pause briefly to motivate the relevance of $b_r(u)$ and $D_2(f)$.
+Suppose one could upper bound $b_r(u)$ by $D_2\bigl(\mc{I}(u)\bigr)$ for an appropriate choice of interpolating map $\mc{I}: \Leb^2(P_n) \to \Leb^2(P)$, and vice versa upper bound $D_2(f)$ by $b_r(\mc{P}(f))$ for an appropriate choice of discretization map $\mc{P}: \Leb^2(P) \to \Leb^2(P_n)$.
+Suppose further that $\mc{I}$ and $\mc{P}$ were near-isometries, meaning $\|\mc{I}(u)\|_{\Leb^2(P)} \approx \|u\|_{\Leb^2(P_n)}$ and $\|\mc{P}(f)\|_{\Leb^2(P_n)} \approx \|f\|_{\Leb^2(P)}$.
+Then using the variational characterization of eigenvalues $\lambda_k(\Delta_P)$ and $\lambda_k(G_{n,r})$---i.e.
+the Courant-Fischer Theorem---one could obtain estimates on the error of $|\lambda_k(\Delta_P) - \lambda_k(G_{n,r})|$, and thereby establish Lemma~\ref{lem:neighborhood_eigenvalue}.
+
+We will momentarily define the maps $\wt{\mc{I}}$ and $\wt{\mc{P}}$, and subsequently prove they satisfy the outlined desiderata.
+These maps will be defined with respect to a particular probability measure $\wt{P}_n$ that with high probability is close in transportation distance to both $P_n$ and $P$---this probabilistic estimate is the workhorse that allows us to relate the random energy $b_r$ to the non-random $D_2$.
+We now review the notion of transportation distance, as well formalize the properties that $\wt{P}_n$ will satisfy.
+
+\paragraph{Transportation distance between $P_n$ and $P$.}
+For a map $T$ defined on $\Xset$ and a measure $\mu$, let $T_{\sharp}\mu$ denote the push-forward of $\mu$ by $T$, i.e the measure for which
+\begin{equation*}
+\bigl(T_{\sharp}\mu\bigr)(U) := \mu\bigl(T^{-1}(U)\bigr)
+\end{equation*}
+for any Borel subset $U \subseteq \Xset$.
+Suppose $T_{\sharp}\mu = P_n$; then the map $T$ is referred to as transportation map between $\mu$ and $P_n$.
+The  $\infty$-transportation distance between $\mu$ and $P_n$ is then
+\begin{equation*}
+d_{\infty}(\mu,P_n) := \inf_{T: T_{\sharp} \mu = P_n} \biggl\{\sup_{x \in \Xset}~\bigl|T(x) - x\bigr|\biggr\}
+\end{equation*}
+\cite{calder2019} consider $\Xset$ a smooth manifold without boundary, i.e.
+they assume $\Xset$ satisfies~\ref{asmp:domain_manifold}.
+They exhibit an absolutely continuous measure $\wt{P}_n$ with density $\wt{p}_n$ which with high probability is close to $P_n$ in transportation distance, for which $\|p - \wt{p}_n\|_{\Leb^\infty}$ is also small.
+In Proposition~\ref{prop:optimal_transport}, we adapt this result to  the setting of full-dimensional manifolds with boundary.
+\begin{proposition}[Restatement of Proposition~2.11 of \textcolor{red}{(Calder2019)}]
+	\label{prop:optimal_transport}
+	Suppose $p$ satisfies~\ref{asmp:bounded_lipschitz_density} and $\Xset$ satisfies~\ref{asmp:domain}.
+Then with probability at least $1 - A_2 n \exp\bigl\{-a_2 n\theta^2\wt{\delta}^d\bigr\}$, the following statement holds: there exists a probability measure $\wt{P}_n$ with density $\wt{p}_n$ such that:
+	\begin{equation*}
+	d_{\infty}(\wt{P}_n, P_n) \leq A_2 \wt{\delta}
+	\end{equation*}
+	and
+	\begin{equation*}
+	\|\wt{p}_n - p\|_{\infty} \leq A_2\bigl(\wt{\delta} + \theta\bigr)
+	\end{equation*}
+\end{proposition}
+ For the rest of this section, we let $\wt{P}_n$ be a probability measure with density $\wt{p}_n$, that satisfies the conclusions of Proposition~\ref{prop:isometry}.
+Additionally we denote by $\wt{T}_n$ an \emph{optimal transport map} between $\wt{P}_n$ and $P_n$.
+Finally, we write $U_1,\ldots,U_n$ for the preimages of $X_1,\ldots,X_n$ under $\wt{T}_n$.
+
+\paragraph{Interpolation and discretization maps.}
+
+The discretization map  $\wt{\mathcal{P}}: \Leb^2(P) \to \Leb^2(P_n)$ is given by
+\begin{equation*}
+\bigl(\wt{\mathcal{P}}f\bigr)(X_i) := n \cdot \int_{U_i} f(x) \wt{p}_n(x) \,dx.
+\end{equation*}
+On the other hand, the interpolation map $\wt{\mc{I}}: \Leb^2(P_n) \to \Leb^2(\Xset)$ is defined as $\wt{\mc{I}}u := \Lambda_{r - 2\wt{\delta}}(\wt{\mc{P}}^{\star}u)$.
+Here, $\wt{\mc{P}}^{\star} = u \circ \wt{T}$ is the adjoint of $\wt{\mc{P}}$, i.e.
+\begin{equation*}
+\bigl(\wt{\mc{P}}^{\star}u\bigr)(x) = \sum_{j = 1}^{n} u(x_i) \1\{x \in U_i\} 
+\end{equation*} 
+and $\Lambda_r$ is a smoothening of this extension, formally
+\begin{equation*}
+\Lambda_r(f) := \frac{1}{r^d\tau(x)}\int_{\Xset} \eta_r(x',x) f(x') \,dx',~~ \eta_r(x',x) := \psi\biggl(\frac{\|x' - x\|}{r}\biggr)
+\end{equation*}
+where $\psi(t) := (1/\sigma_K)\int_{t}^{\infty} s K(s) \,ds$ and $\tau(x) := (1/r^d)\int_{\Xset} \eta_r(x',x) \,dx'$ is a normalizing constant.
+
+Propositions~\ref{prop:dirichlet_energies} and~\ref{prop:isometry} establish our claims regarding $\wt{\mc{P}}$ and $\wt{\mc{I}}$: first, that they approximately preserve the Dirichlet energies $b_r$ and $D_2$, and second that they are near-isometries for functions $u \in \Leb^2(P_n)$ (or $f \in \Leb^2(P)$) of small Dirichlet energy $b_r(u)$ (or $D_2(f)$).
+
+\begin{proposition}[\textbf{c.f. Proposition 4.1 of \citet{calder2019}}]
+	\label{prop:dirichlet_energies}
+	With probability at least $1 - A_2n\exp(-a_2n\theta^2\wt{\delta}^{m})$, we have the following.
+	\begin{enumerate}[(1)]
+		\item For any $u \in \Leb^2(P_n)$,
+		\begin{equation*}
+		\sigma_{K} D_2(\wt{\mc{I}}u) \leq \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) b_r(u)
+		\end{equation*}
+		\item For any $f \in \Leb^2(\Xset)$,
+		\begin{equation*}
+		b_r(\wt{\mc{P}}f) \leq \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + L_p(2r + 4\wt{\delta})\Bigr) \cdot \Bigl(1 + A_4\frac{\wt{\delta}}{r}\Bigr) \sigma_{K} D_2(f).
+		\end{equation*}
+	\end{enumerate}
+\end{proposition}
+
+\begin{proposition}[\textbf{c.f. Proposition 4.2 of \citet{calder2019}}]
+	\label{prop:isometry}
+	With probability at least $1 - A_2n\exp(-a_2n\theta^2\wt{\delta}^{m})$, we have the following.
+	\begin{enumerate}[(1)]
+		\item For every $f \in \Leb^2(\Xset)$,
+		\begin{equation*}
+		\Bigl|\|f\|_{\Leb^2(P)}^2 - \|\wt{P}f\|_{\Leb^2(P_n)}^2\Bigr| \leq A_5 \wt{\delta} \|f\|_{\Leb^2(P)} \sqrt{D_2(f)} + \frac{A_1\bigl(\theta + \wt{\delta}\bigr)}{p_{\min}} \|f\|_{\Leb^2(P)}^2
+		\end{equation*}
+		\item For every $u \in \Leb^2(P_n)$,
+		\begin{equation*}
+		\Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(P)}^2 - \|u\|_{\Leb^2(P_n)}^2\Bigr| \leq C r \|u\|_{\Leb^2(P_n)} \sqrt{b_r(u)} + C\bigl(\theta + \wt{\delta}\bigr) \|u\|_{\Leb^2(P_n)}^2
+		\end{equation*}
+	\end{enumerate}
+\end{proposition}
+
+Given these propositions, the proof of Lemma~\ref{lem:neighborhood_eigenvalue} follows by the Courant-Fischer Theorem.
+
+\paragraph{Proof of Lemma~\ref{lem:neighborhood_eigenvalue}.}
+We start with the upper bound, proceeding exactly as in Proposition 4.4 of \citep{burago2014}.
+Let $W$ be the span of $f_1,\ldots,f_{k}$, the first $k$ eigenfunctions of $-\Delta_P$, so that by the Courant-Fischer principle $D_2(f) \leq \lambda_k(\Delta_P) \|f\|_{\Leb^2(P)}$ for every $f \in W$.
+As a result, by Part (1) of Proposition~\ref{prop:isometry} we have that for any $f \in W$
+\begin{equation*}
+\bigl\|\wt{\mc{P}}f\bigr\|_{\Leb^2(P_n)}^2 \geq \biggl(1 - A_5\wt{\delta} \sqrt{\lambda_{k}(\Delta_P)} - \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr)\|f\|_{\Leb^2(P)}^2  \geq \frac{1}{2} \|f\|_{\Leb^2(P)}^2,
+\end{equation*}
+where the second inequality follows by assumption~\eqref{eqn:neighborhood_eigenvalue_1}.
+
+Therefore $\wt{\mc{P}}$ is injective over $W$, and $\wt{\mc{P}}W$ has dimension $\ell$.
+This means we can invoke the Courant-Fischer Theorem, along with Proposition \ref{prop:dirichlet_energies}, and conclude that
+\begin{align*}
+\lambda_k(G_{n,r}) & \leq \max_{\substack{u \in \wt{\mc{P}}W \\ u \neq 0} } \frac{b_r(u)}{\|u\|_{\Leb^2(P_n)}^2} \\
+& = \max_{\substack{f \in W \\ f \neq 0} } \frac{b_r(\wt{\mc{P}}f)}{\bigl\|\wt{\mc{P}}f\bigr\|_{\Leb^2(P_n)}^2} \\
+& \leq \frac{\Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + L_p(2r + 4\wt{\delta})\Bigr) \cdot \Bigl(1 + \frac{2L_K}{K(1/2)}\frac{\wt{\delta}}{r}\Bigr)}{1 - A_5\wt{\delta} \sqrt{\lambda_{\ell}(\Delta_P)} - 2\frac{A_1(\theta + \wt{\delta})}{p_{\min}}} \sigma_K \lambda_{\ell}(\Delta_P),
+\end{align*}
+establishing the upper bound in~\eqref{eqn:eigenvalue_bound} upon proper choice of $A$.
+
+\paragraph{Organization of this section.}
+The rest of this section will be devoted to proving Propositions~\ref{prop:optimal_transport},~\ref{prop:dirichlet_energies} and~\ref{prop:isometry}.
+The proof of Proposition~\ref{prop:optimal_transport} is given in Subsection~\ref{subsec:proof_proposition_optimal_transport}.
+To prove the latter two propositions, it will help to introduce the intermediate energies
+\begin{equation*}
+\wt{E}_r(f,\eta,V) := \frac{1}{r^{d + 2}}\int_{V} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{r}\biggr) \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx
+\end{equation*}
+and
+\begin{equation*}
+{E}_r(f,\eta,V) := \frac{1}{r^{d + 2}}\int_{V} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx
+\end{equation*}
+where $\eta: [0,\infty) \to [0,\infty)$ is an arbitrary kernel, and $V \subseteq \Xset$ is a measurable set.
+We will abbreviate $\wt{E}_r(f,\eta,\Xset)$ as $\wt{E}_r(f,\eta)$ and $\wt{E}_r(f,K) = \wt{E}_r(f)$ (and likewise with $E_r$.)
+
+In Subsection~\ref{subsec:integrals}, we establish relationships between the (non-random) functionals $E_r(f)$ and $D_2(f)$, as well as providing estimates on some assorted integrals.
+In Subsection~\ref{subsec:random_functionals}, we establish relationships between the stochastic functionals $\wt{E}_r(f)$ and $E_r(f)$,  between $\wt{E}_r\bigl(\wt{\mc{I}}(u)\bigr)$ and $b_r\bigl(u\bigr)$, and between $\wt{E}_r\bigl(f\bigr)$ and $b_r\bigl(\wt{\mc{P}}f\bigr)$.
+Finally, in Subsection~\ref{subsec:proof_of_prop_dirichlet_energies_and_isometry} we use these various relationships to prove Propositions~\ref{prop:dirichlet_energies} and~\ref{prop:isometry}.
+
+\subsection{Proof of Proposition~\ref{prop:optimal_transport}}
+\label{subsec:proof_proposition_optimal_transport}
+
+We start by defining the density $\wt{p}_n$, which will be piecewise constant over a particular partition $\mc{Q}$ of $\Xset$; specifically if $x \in Q$ for $Q \in \mc{Q}$, then
+\begin{equation}
+\label{pf:prop_optimal_transport_0}
+\wt{p}_n(x) := \frac{P_n(Q)}{\vol(Q)}
+\end{equation}
+where $\vol(\cdot)$ denotes the full-dimensional Lebesgue measure.
+
+We now construct the partition $\mc{Q}$, in progressive degrees of generality on the domain $\Xset.$
+\begin{itemize}
+	\item In the special case of the unit cube $\Xset = (0,1)^d$, the partition will simply be the collection of cubes
+	\begin{equation*}
+	\set{Q_k: k \in [\wt{\delta}^{-1}]^d} 
+	\end{equation*}
+	where $Q_k = \wt{\delta}\Bigl([k_1 - 1,k_1] \otimes \cdots \otimes [k_d - 1,k_d]\Bigr)$ and we assume without loss of generality that $\wt{\delta}^{-1} \in \mathbb{N}$.
+	\item If $\mc{\Xset}$ is an open, connected set with smooth boundary, then by Proposition 3.2 of \textcolor{red}{GarciaTrillos14}, there exist a finite number $N(\Xset) \in \mathbb{N}$ of disjoint polytopes which cover $\Xset$.
+Moreover, letting $U_j$ denote the intersection of the $j$th of these polytopes with $\wb{\Xset}$, this proposition establishes that for each $j$ there exists a bi-Lipschitz homeomorphism $\Phi_j: U_j \to [0,1]^d$.
+We take the collection of
+	\begin{equation*}
+	\mc{Q} = \biggl\{\Phi_j^{-1}(Q_k): j = 1,\ldots,N(\Xset)~~\textrm{and}~~k \in [\wt{\delta}^{-1}]^d\biggr\}
+	\end{equation*}
+	to be our partition. Denote by $L_{\Phi}$ the maximum of the bi-Lipschitz constants of $\Phi_1,\ldots,\Phi_{N(\Xset)}$.
+	\item Finally, in the general case where $\Xset$ is an open, connected set with Lipschitz boundary, then there exists a bi-Lipschitz homeomorphism $\Psi$ between $\Xset$ and a smooth, open, connected set with Lipschitz boundary.
+Letting $\Phi_j$ and $\wt{Q}_{j,k}$ be as before, we take the collection
+	\begin{equation*}
+	\mc{Q} = \biggl\{\wt{Q}_{j,k} = \Bigl(\Psi^{-1} \circ \Phi_j^{-1}\Bigr)(Q_k): j = 1,\ldots,N(\Xset)~~\textrm{and}~~k \in [\wt{\delta}^{-1}]^d\biggr\}
+	\end{equation*}
+	to be our partition. Denote by $L_{\Psi}$ the bi-Lipschitz constant of $\Psi$.
+\end{itemize}
+Let us record a few facts which hold for all $\wt{Q}_{j,k} \in \mc{Q}$, which follow from the bi-Lipschitz properties of $\Phi_j$ and $\Psi$: first that
+\begin{equation}
+\label{pf:prop_optimal_transport_1}
+\diam(\wt{Q}_{j,k}) \leq L_{\Psi} \L_{\Phi} \wt{\delta}
+\end{equation}
+and second that
+\begin{equation}
+\label{pf:prop_optimal_transport_2}
+\vol(\wt{Q}_{j,k}) \geq \biggl(\frac{1}{L_{\Psi} L_{\Phi}}\biggr)^d \wt{\delta}^d.
+\end{equation}
+We now use these facts to show that $\wt{P}_n$ satisfies the claims of Proposition~\ref{prop:optimal_transport}.
+One on the one hand, by the construction~\eqref{pf:prop_optimal_transport_0} clearly there exists a transport map $\wt{T}: \Xset \to \mathbf{X}$ for which every $x \in \wt{Q}_{j,k}$ is moved to a design point $X \in {\bf X} \cap \wt{Q}_{j,k}$, and so by~\eqref{pf:prop_optimal_transport_1}
+\begin{equation*}
+\sup_{x \in \Xset} \abs{\wt{T}(x) - x} \leq  L_{\Psi} \L_{\Phi} \wt{\delta}.
+\end{equation*}
+On the other hand, applying the triangle inequality we have that for $x \in \wt{Q}_{j,k}$
+\begin{align*}
+|\wt{p}_n(x) - p(x)| \leq \biggl|\frac{P_n(\wt{Q}_{j,k}) - P(\wt{Q}_{j,k})}{\vol(\wt{Q}_{j,k})}\biggr| + \frac{1}{\vol(\wt{Q}_{j,k})} \int_{\wt{Q}_{j,k}} |p(x') - p(x)| \,dx 
+\end{align*}
+and using the Lipschitz property of $p$ we find that 
+\begin{equation}
+\label{pf:prop_optimal_transport_3}
+\|\wt{p}_n - p\|_{\Leb^{\infty}} \leq \max_{j,k} \biggl|\frac{P_n(\wt{Q}_{j,k}) - P(\wt{Q}_{j,k})}{\vol(\wt{Q}_{j,k})}\biggr| + L_p L_{\Phi} L_{\Psi} \wt{\delta}
+\end{equation}
+From Hoeffding's inequality and a union bound, we obtain that 
+\begin{align*}
+\mathbb{P}\biggl( \bigl|P_n(\wt{Q}) - P(\wt{Q})\bigr| & \leq \theta P(\wt{Q}) \quad \forall \wt{Q} \in \mc{Q} \biggr) \geq 1 - 2 \sharp(\mc{Q}) \cdot \exp\biggl\{-\frac{\theta^2 n \min \{P(\wt{Q})\}}{3}\biggr\} \\
+& \geq 1 - \frac{2 N(\Xset)}{\wt{\delta}^d} \cdot \exp\biggl\{-\frac{\theta^2 n p_{\min} \wt{\delta}^d }{3\bigl(L_{\Psi} L_{\Phi}\bigr)^d}\biggr\}.
+\end{align*}
+Noting that by assumption $P(\wt{Q}) \leq p_{\max} \vol(\wt{Q})$ and $\wt{\delta}^{-d} \leq n$, the claim follows upon plugging back into~\eqref{pf:prop_optimal_transport_3}, and setting
+\begin{equation*}
+a_1(d, \Xset) := \frac{1}{3\bigl(L_{\Psi} L_{\Phi}\bigr)^d}~~\textrm{and}~~A_1(d,\Xset) := \max \Bigl\{2N(\Xset),L_p L_{\Psi} L_{\Phi} \Bigr\}
+\end{equation*}
+in the statement of the proposition.
+
+\subsection{Non-random functionals and integrals}
+\label{subsec:integrals}
+
+Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected} relates the non-local energy $E_r(f)$ to $D_2(f)$.
+It follows in large part from the proofs of Lemma 13 of \cite{trillos2019} and Lemma 6 of \cite{burago2014}, with minor modifications to account for the boundary of $\Xset$.
+\begin{lemma}[c.f. Lemma 13 of \cite{trillos2019}, Lemma 6 of \cite{burago2014}]
+	\label{lem:first_order_graph_sobolev_seminorm_expected}
+	For any $f \in \Leb^2(\Xset)$,
+	\begin{equation*}
+	E_r(f) \leq (1 + L_pr)^2 \cdot \sigma_K D_2(f)
+	\end{equation*}
+\end{lemma}
+
+In Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, we establish the reverse of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected}. 
+\begin{lemma}[c.f. Lemma 9 of \cite{trillos2019}, Lemma 5.5 of \cite{burago2014}]
+	\label{lem:first_order_graph_sobolev_seminorm_expected_lb}
+	For any $f \in \Leb^2(\Xset)$,
+	\begin{equation*}
+	\sigma_KD_2(\Lambda_rf) \leq \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K}E_r(f)
+	\end{equation*}
+\end{lemma}
+
+To prove Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, we require upper and lower bounds on $\tau(x)$, as well as an upper bound on the gradient of $\tau$.
+The lower bound here---$\tau(x) \geq 1/2$---is quite a bit a looser than in the case where $\Xset$ has no boundary---$\tau(x) \geq (1 + Cr^2)^{-1}$.
+The same is the case regarding the upper bound of the size of the gradient $\|\nabla \tau(x)\|$.
+However, the bounds as stated here will be sufficient for our purposes.
+\begin{lemma}
+	\label{lem:tau_bound}
+	Suppose $r$ satisfies~\ref{asmp:r_small_3}. For all $x \in \Xset$,
+	\begin{equation*}
+	\frac{1}{2} \leq \tau(x) \leq 1,
+	\end{equation*}
+	and additionally $\|\nabla \tau(x)\| \leq \frac{1}{\sigma_K r}$.
+\end{lemma}
+
+Finally, to prove part (2) of Proposition~\ref{prop:isometry}, we require Lemma~\ref{lem:smoothening_error}, which gives an estimate on the error $\Lambda_h f - f$ in $\Leb^2(P)$ norm.
+\begin{lemma}[c.f Lemma 8 of \cite{trillos2019}, Lemma 5.4 of \cite{burago2014}]
+	\label{lem:smoothening_error}
+	For any $h > 0$, 
+	\begin{equation}
+	\label{eqn:smoothening_error_norm}
+	\bigl\|\Lambda_hf\bigr\|_{\Leb^2(P)}^2 \leq \frac{2p_{\max}}{p_{\min}} \bigl\|f\bigr\|_{\Leb^2(P)}^2
+	\end{equation}
+	and
+	\begin{equation}
+	\label{eqn:smoothening_error_energy}
+	\bigl\|\Lambda_hf - f\bigr\|_{\Leb^2(P)}^2 \leq \frac{2}{\sigma_Kp_{\min}} r E_r(f)
+	\end{equation}
+	for all $f \in \Leb^2(\Xset)$.
+\end{lemma}
+
+\paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected}.}
+Suppose $f \in C^{\infty}(\wb{\Xset})$, which we may do without loss of generality due to the density of $C^{\infty}(\wb{\Xset})$ in  $\Leb^2(\Xset)$.
+
+Taylor expanding $f(x')$ about $x' = x$, applying first Cauchy-Schwarz' and then Jensen's inequality, and then using the Lipschitz property of $p$, we have
+\begin{align*}
+E_r(f) & = \frac{1}{r^{d + 2}}\int_{\Xset} \int_{\Xset} (f(x') - f(x))^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \\
+& = \frac{1}{r^{d + 2}}\int_{\Xset} \int_{\Xset} \biggl(\sum_{i = 1}^{d} (x' - x)^{e_i} \int_0^1 f^{(e_i)}\bigl(x + t(x' - x)\bigr) \,dt\biggr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \\
+& \leq \frac{1}{r^{d + 2}}\int_{\Xset} \int_{\Xset} \|x' - x\|^2 \sum_{i = 1}^{d}\biggl(\int_0^1 f^{(e_i)}\bigl(x + t(x' - x)\bigr) \,dt\biggr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \\
+& \leq \frac{1}{r^{d + 2}}\int_{\Xset} \int_{\Xset} \int_0^1   \|x' - x\|^2 \biggl\{\sum_{i = 1}^{d} \Bigl(f^{(e_i)}\bigl(x + t(x' - x)\bigr)\Bigr)^2\biggr\} K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x)   \,dt   \,dx' \,dx \\
+& \leq \frac{(1 + L_pr)^2}{r^{d + 2}}\int_{\Xset} \int_{\Xset} \int_0^1  \|x' - x\|^2 \biggl\{\sum_{i = 1}^{d} \Bigl(f^{(e_i)}\bigl(x + t(x' - x)\bigr)\Bigr)^2\biggr\} K\biggl(\frac{\|x' - x\|}{r}\biggr) p\bigl(x + t(x' - x)\bigr)^2   \,dt   \,dx' \,dx.
+\end{align*} 
+Then letting $z = (x' - x)/r$ and $\wt{x} = x + trz$, we obtain
+\begin{align*}
+E_r(f) & \leq (1 + L_pr)^2 \int_{\Xset} \int_{\mathcal{Z}_r(x)} \int_{0}^{1} \|z\|^2 \biggl\{\sum_{i = 1}^{d} \Bigl(f^{(e_i)}\bigl(x + trz\bigr)\Bigr)^2\biggr\} K\bigl(\|z\|\bigr) p(x + trz)^2 \,dt \,dz \,dx \\
+& = (1 + L_pr)^2 \int_{B(0,1)} \int_{0}^{1}  \int_{\Xset_r(z)} \|z\|^2 \biggl\{\sum_{i = 1}^{d} \Bigl(f^{(e_i)}\bigl(x + trz\bigr)\Bigr)^2\biggr\} K\bigl(\|z\|\bigr) p(x + trz)^2 \,dx \,dt \,dz \\
+& \leq (1 + L_pr)^2 \int_{B(0,1)} \int_{0}^{1} \|z\|^2 K\bigl(\|z\|\bigr) \biggl\{\int_{\wt{\Xset}_r(t,z)} \|\nabla f(\wt{x})\|^2 p(\wt{x})^2 \,d\wt{x}\biggr\} \,dt \,dz  \\
+& \leq (1 + L_pr)^2 \int_{B(0,1)} \int_{0}^{1} \|z\|^2 K\bigl(\|z\|\bigr) \biggl\{\int_{\Xset} \|\nabla f(\wt{x})\|^2 p(\wt{x})^2 \,d\wt{x}\biggr\} \,dt \,dz \\
+& = (1 + L_pr)^2 \int_{B(0,1)} \|z\|^2 K\bigl(\|z\|\bigr) \,dz 
+\end{align*}
+where $\mathcal{Z}_r(x) = B(0,1) \cap (\Xset - x)/r$, $\Xset_r(z) = \Xset \cap( \Xset - zr)$, $\wt{\Xset}_r(t,z) = \Xset_r(z) + trz$, and we note that since
+\begin{align*}
+\wt{x} \in \wt{\Xset}_r(t,z) & \Longrightarrow \wt{x} = x + trz ~~ \textrm{for some $x \in \Xset - zr$ and $t \in [0,1]$} \\
+& \Longrightarrow \wt{x} \in \Xset,~~ \textrm{(since $x \in \Xset$, $x + rz \in \Xset$ and $\Xset$ is connected)}
+\end{align*}
+it must be that $\wt{X}_r(t,z) \subseteq \Xset$ for all $z \in B(0,1)$ and $t \in [0,1]$.
+
+\paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}.}
+For any $a \in \Reals$, $\Lambda_rf$ satisfies the identity
+\begin{equation*}
+\Lambda_rf(x) = a + \frac{1}{r^d\tau(x)}\int_{\Xset} \eta_r(x',x)\bigl(f(x') - a\bigr)\,dx'
+\end{equation*}
+and by differentiating we obtain
+\begin{equation*}
+\bigl(\nabla \Lambda_rf\bigr)(x)= \frac{1}{r^d\tau(x)}\int_{\Xset} \bigl(\nabla \eta_r(x',\cdot)\bigr)(x)\bigl(f(x') - a\bigr)\,dx' + \nabla\bigl(\tau^{-1}\bigr)(x)\cdot \frac{1}{r^d}\int_{\Xset} \eta_r(x',x)\bigl(f(x') - a\bigr)\,dx'
+\end{equation*} 
+Plugging in $a = f(x)$, we get $\nabla\Lambda_rf(x) = J_1(x)/\tau(x) + J_2(x)$ for
+\begin{equation*}
+J_1(x) := \frac{1}{r^d}\int_{\Xset} \bigl(\nabla \eta_r(x',\cdot)\bigr)(x)\bigl(f(x') - f(x)\bigr)\,dx',~~ J_2(x) := \nabla\bigl(\tau^{-1}\bigr)(x)\cdot \frac{1}{r^d}\int_{\Xset} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)\,dx'.
+\end{equation*}
+To upper bound $\bigl\|J_1(x)\bigr\|^2$, we first compute the gradient of $\eta_r(x',\cdot)$,
+\begin{align*}
+\bigl(\nabla\eta_r(x',\cdot)\bigr)(x) & = \frac{1}{r} \psi'\biggl(\frac{\|x'  - x\|}{r}\biggr) \frac{(x - x')}{\|x' - x\|} \\
+& = \frac{1}{\sigma_Kr^2} K\biggl(\frac{\|x' - x\|}{r}\biggr) (x' - x),
+\end{align*}
+so that by the Cauchy-Schwarz inequality,
+\begin{align*}
+\bigl\|J_1(x)\bigr\|^2 & = \frac{1}{\sigma_K^2 r^{4 + 2d}} \Biggl[\int_{\Xset} \bigl(f(x') - f(x)\bigr)K\biggl(\frac{\|x' - x\|}{r}\biggr)(x' - x)\Biggr]^2 \,dx' \\
+& \leq \frac{1}{\sigma_K^2r^{4 + 2d}} \biggl[\int_{\Xset}K\biggl(\frac{\|x' - x\|}{r}\biggr)\|x' - x\|^2\,dx'\biggr] \biggl[\int_{\Xset}K\biggl(\frac{\|x' - x\|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2\,dx'\biggr] \\
+& \leq \frac{1}{\sigma_K r^{2 + d}}\int_{\Xset}K\biggl(\frac{\|x' - x\|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2\,dx'.
+\end{align*}
+To upper bound $\bigl\|J_2(x)\bigr\|^2$, we use the Cauchy-Schwarz inequality along with the observation $\eta_r(x',x) \leq \frac{1}{\sigma_K} K\bigl(\|x' - x\|/r\bigr)$ to deduce
+\begin{align*}
+\bigl\|J_2(x)\bigr\|^2 & \leq \Bigl\|\nabla\bigl(\tau^{-1}\bigr)(x)\Bigr\|^2\frac{1}{r^{2d}} \biggl[\int_{\Xset}\eta_r(x',x) \,dx'\biggr] \cdot \biggl[\int_{\Xset} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr] \\
+& = \Bigl\|\nabla\bigl(\tau^{-1}\bigr)(x)\Bigr\|^2\frac{\tau(x)}{r^d} \int_{\Xset} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)^2 \,dx'\\ 
+& \leq \Bigl\|\nabla\bigl(\tau^{-1}\bigr)(x)\Bigr\|^2\frac{\tau(x)}{\sigma_K r^d}\int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx' \\
+& \leq \frac{16}{\sigma_K^2r^{2 + d}} \biggl[\int_{\Xset} K\biggl(\frac{|x' - x|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr].
+\end{align*}
+where the last inequality follows by the estimates on $\tau$ and $\nabla \tau$ given in Lemma~\ref{lem:tau_bound}.
+Combining our bounds on $\bigl\|J_1(x)\bigr\|^2$ and $\bigl\|J_2(x)\bigr\|^2$ along with the lower bound on $\tau(x)$ in Lemma~\ref{lem:tau_bound} and integrating over $\Xset$, we have
+\begin{align*}
+\sigma_K D_2(\Lambda_r f) & = \sigma_K\int_{\Xset} \biggl\|\Bigl(\nabla \Lambda_rf)(x)\biggr\|^2 p^2(x) \,dx \\
+& \leq 2 \sigma_K \int_{\Xset} \Biggl(\frac{\bigl\|J_1(x)\bigr\|^2}{\tau^2(x)} + \bigl\|J_2(x)\bigr\|^2\Biggr) p^2(x) \,dx \\
+& \leq \frac{40}{\sigma_K r^{2 + d}} \int_{\Xset} \int_{\Xset} K\biggl(\frac{|x' - x|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 p^2(x) \,dx' \,dx \\
+& \leq \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} E_r(f),
+\end{align*}
+completing the proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}.
+
+\paragraph{Proof of Lemma~\ref{lem:tau_bound}.}
+First prove the estimates of $\tau(x)$, and then upper bound on $\|\nabla\tau(x)\|$. Substituting $z = (x' - x)/r$ and using~\ref{asmp:r_small_3}, we have that
+\begin{align*}
+\tau(x) & = \frac{1}{r^d} \int_{\Xset} \psi\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx' \\
+& =  \int_{\mc{B}_x} \psi\bigl(\|z\|\bigr) \,dz \\
+& \geq a_1(\Xset,d) \cdot \int_{B(0,1)} \psi(\|z\|) \,dz
+\end{align*}
+We will now show that $\int_{B(0,1)} \psi(\|z\|) \,dz = 1$, which will imply our estimate of $\tau(x)$.
+To see this identity, note that on the one hand, by converting to polar coordinates and integrating by parts we obtain
+\begin{align*}
+\int_{B(0,1)} \psi\bigl(\|z\|\bigr) \,dz = d \nu_d \int_{0}^{1} \psi(t) t^{d - 1} \,dt = \nu_d \int_{0}^{1} \psi'(t) t^{d} \,dt = \frac{\nu_d}{\sigma_K} \int_{0}^{1} t^{d + 1} K(t) \,dt;
+\end{align*}
+on the other hand, again converting to polar coordinate, we have
+\begin{equation*}
+\sigma_K = \frac{1}{d} \int_{\Reals^d} \|x\|^2 K(\|x\|) \,dx = \nu_d \int_{0}^{1}t^{d + 1} K(t) \,dt.
+\end{equation*}
+
+Finally, we upper bound $\|\nabla\tau(x)\|^2$. Exchanging derivative and integral, we have
+\begin{align*}
+\nabla\tau(x) = \frac{1}{r^d} \int_{\Xset} \bigl(\nabla \eta_r(x',\cdot)\bigr)(x) \,dx' = \frac{1}{\sigma_K r^{d + 2}} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr)(x' - x)\,dx',
+\end{align*}
+and by the Cauchy-Schwarz inequality,
+\begin{equation*}
+\|\nabla\tau(x)\|^2 \leq \frac{1}{\sigma_K^2 r^{2d + 4}} \biggl[\int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr)\,dx'\biggr] \int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr)\|x' - x\|^2\,dx', \leq \frac{1}{\sigma_K r^{2}}
+\end{equation*}
+concluding the proof of Lemma~\ref{lem:tau_bound}. One note: while $\nabla\tau(x) = 0$ when $B(x,r) \in \Xset$, near the boundary the upper bound we derived by using Cauchy-Schwarz appears tight.
+
+\paragraph{Proof of Lemma~\ref{lem:smoothening_error}.}
+By Jensen's inequality and Lemma~\ref{lem:tau_bound},
+\begin{align*}
+\Bigl[\Lambda_rf(x)\Bigr]^2 & \leq \frac{1}{r^d \tau(x)}\int_{\Xset} \eta_r(x',x) \bigl[f(x')\bigr]^2 \,dx' \\
+& \leq \frac{2}{r^d p_{\min}}\int_{\Xset} \eta_r(x',x) \bigl[f(x')\bigr]^2 p(x') \,dx'
+\end{align*}
+Then integrating over $x$ and swapping the order of integraton, we have
+\begin{align*}
+\bigl\|\Lambda_rf\bigr\|_{\Leb^2(P)}^2 & \leq \frac{2}{r^d p_{\min}} \int_{\Xset} \int_{\Xset} \eta_r(x',x) \bigl[f(x')\bigr]^2 p(x') p(x) \,dx' \,dx \\ 
+& \leq \frac{2p_{\max}}{p_{\min}} \int_{\Xset} \bigl[f(x')\bigr]^2 p(x') \,dx' \\
+& = \frac{2p_{\max}}{p_{\min}} \|f\|_{\Leb^2(P)}^2.
+\end{align*}
+
+Noting that $\Lambda_ra = a$ for any $a \in \Reals$, by the Cauchy-Schwarz inequality we have that for all $x \in \Xset$,
+\begin{align*}
+\bigl|\Lambda_rf(x) - f(x)\bigr|^2 & = \biggl[\frac{1}{r^d\tau(x)} \int_{\Xset} \eta_r(x',x) \bigl(f(x') - f(x)\bigr) \,dx'\biggr]^2 \\
+& \leq \frac{1}{r^{2d} \tau^2(x)} \biggl[\int_{\Xset} \eta_r(x',x) \,dx'\biggr] \cdot \biggl[\int_{\Xset} \eta_r(x',x) \bigl(f(x') - f(x)\bigr)^2 \,dx'\biggr] \\
+& = \frac{1}{r^d \tau(x)} \int_{\Xset} \eta_r(x',x) \bigl(f(x') - f(x)\bigr)^2 \,dx'. \\
+& \leq \frac{1}{r^d \tau(x) p_{\min}} \int_{\Xset} \eta_r(x',x) \bigl(f(x') - f(x)\bigr)^2 p(x') \,dx'.
+\end{align*}
+Then integrating over $\Xset$ with respect to $p$ yields~\eqref{eqn:smoothening_error_energy}.
+
+\subsection{Random functionals}
+\label{subsec:random_functionals}
+
+We begin by relating $\wt{E}_r(f)$ and $E_r(f)$.
+Note that by assumption $A_2(\wt{\delta} + \theta)/p_{\min} \leq A_1(\wt{\delta} + \theta) \leq \frac{1}{2}$.
+Therefore, some standard calculations show that,
+\begin{equation}
+\label{eqn:calder19_1}
+\bigl(1 - A_1(\theta + \wt{\delta})\bigr) E_r(f) \leq \wt{E}_r(f) \leq \bigl(1 + A_1(\theta + \wt{\delta})\bigr) E_r(f),
+\end{equation}
+as well as implying that the norms $\|f\|_{\Leb^2(P)}$ and $\|f\|_{\Leb^2(P_n)}$ satisfy
+\begin{equation}
+\label{eqn:calder19_2}
+\bigl(1 - A_1(\theta + \wt{\delta})\bigr) \|f\|_{\Leb^2(P)}^2 \leq \|f\|_{\Leb^2(\wt{P}_n)}^2 \leq \bigl(1 + A_1(\theta + \wt{\delta})\bigr) \|f\|_{\Leb^2(P)}^2.
+\end{equation}
+
+Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized} relates the graph Sobolev semi-norm $b_r(\wt{\mc{P}}f)$ to the non-local energy $\wt{E}_r(f)$.
+It follows exactly from the derivations in the proof of Lemma 13 in \cite{trillos2019}; indeed, in our flat Euclidean setting, the proof is simplified, and we include this simpler proof for completeness purposes only.
+\begin{lemma}[\textbf{c.f. Lemma 13 of \cite{trillos2019}, Lemma 4.3 of \cite{burago2014}}]
+	\label{lem:first_order_graph_sobolev_seminorm_discretized}
+	Suppose~\ref{asmp:kernel} is satisfied, and that the conclusions of Proposition~\ref{prop:optimal_transport} are met. Then
+	\begin{equation*}
+	b_r(\wt{\mc{P}}f) \leq \wt{E}_{r + 2\wt{\delta}}(f) + 2\frac{L_K\wt{\delta}}{K(1/2)r} \wt{E}_{2(r + 2\wt{\delta})}(f)
+	\end{equation*}
+	for any $f \in \Leb^2(\Xset)$.
+\end{lemma}
+
+In Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}, we establish the reverse of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}. 
+\begin{lemma}[c.f. Lemma 14 of \cite{trillos2019}]
+	\label{lem:first_order_graph_sobolev_seminorm_discretized_lb}
+	Suppose $\frac{\wt{\delta}}{r} \leq \min\bigl\{\frac{1}{4(2^{d + 2} - 1)}, \frac{K(1)}{2L_K}\bigr\}$. Then for any $u \in \Leb^2(P_n)$, 
+	\begin{equation*}
+	\wt{E}_{r - 2\wt{\delta}}\bigl(\wt{\mc{P}}^{\star}u\bigr) \leq \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) b_{r}(u)
+	\end{equation*}
+\end{lemma}
+
+\paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}.}
+Recalling that $\bigl(\wt{P}f\bigr)(X_i) = n \cdot \int_{U_i} f(x) \wt{p}_n(x)$, the key point is that
+\begin{equation*}
+\biggl(\bigl(\wt{P}f\bigr)(X_i) - \bigl(\wt{P}f\bigr)(X_j)\biggr)^2 \leq n^2 \cdot \int_{U_i} \int_{U_j} \bigl(f(x') - f(x)\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx.
+\end{equation*}
+Additionally, the non-increasing and Lipschitz properties of $K$ imply that for any $x \in U_i$ and $x' \in U_j$, 
+\begin{equation*}
+K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) \leq K\biggl(\frac{\bigl(\|x' - x\| - 2\wt{\delta}\bigr)_{+}}{r}\biggr) \leq K\biggl(\frac{\|x' - x\|}{r + 2\wt{\delta}}\biggr) + \frac{2L_K\wt{\delta}}{r}\1\Bigl\{\|x' - x\| \leq r + 2\wt{\delta}\Bigr\}
+\end{equation*}
+Using this along with the Lipschitz property of $K$, we get
+\begin{align*}
+b_r(f) & = \frac{1}{n^2r^{d + 2}} \sum_{i,j = 1}^n \Bigl(\bigl(\wt{P}f\bigr)(X_i) - \bigl(\wt{P}f\bigr)(X_j)\Bigr)^2 K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) \\
+& \leq \frac{1}{r^{d + 2}} \sum_{i,j = 1}^n \int_{U_i} \int_{U_j}  \bigl(f(x') - f(x)\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) \,dx' \,dx \\
+& \leq \frac{1}{r^{d + 2}} \sum_{i,j = 1}^n \int_{U_i} \int_{U_j}  \bigl(f(x') - f(x)\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) \biggl[K\biggl(\frac{\|x' - x\|}{r + 2\wt{\delta}}\biggr) + \frac{2L_K\wt{\delta}}{r}\1\Bigl\{\|x' - x\| \leq r + 2\wt{\delta}\Bigr\}\biggr] \,dx' \,dx \\
+& = \wt{E}_{r + 2\wt{\delta}}(f) + \frac{2L_K\wt{\delta}}{r}\wt{E}_{r + 2\wt{\delta}}(f; \1_{[0,1]})
+\end{align*}
+for $\1_{[0,1]}(t) = \1\{0 \leq t \leq 1\}$. But clearly $\wt{E}_{r + 2\wt\delta}(f; \1_{[0,1]}) \leq 1/(K(1/2))\wt{E}_{2r + 4\wt{\delta}}(f)$, and the Lemma is shown.
+
+\paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}.}
+For brevity, we write $\wt{r} := r - 2\wt{\delta}$. We begin by expanding the energy $\wt{E}_{r - 2\wt{\delta}}\bigl(\wt{\mc{P}}^{\star}u\bigr)$ as a double sum of double integrals,
+\begin{align*}
+\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & = \frac{1}{\wt{r}^{d + 2}} \sum_{i = 1}^{n} \sum_{j = 1}^{n} \int_{U_i} \int_{U_j} \Bigl(u(X_i) - u(X_j)\Bigr)^2 K\biggl(\frac{\|x' - x\|}{\wt{r}}\biggr) \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx
+\end{align*}
+We next use the Lipschitz property of the kernel $K$---in particular that for $x \in U_i$ and $x' \in U_j$,
+\begin{equation*}
+K\biggl(\frac{\|x' - x\|}{\wt{r}}\biggr) \leq K\biggl(\frac{\|x_i - x_j\|}{r}\biggr) + \frac{2L_K\wt{\delta}}{\wt{r}} \cdot \1\biggl\{\frac{\|x' - x\|}{\wt{r}} \leq 1\biggr\}
+\end{equation*}
+---to conclude that
+\begin{align}
+\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & \leq \frac{1}{n^2\wt{r}^{d + 2}} \sum_{i = 1}^{n} \sum_{j = 1}^{n} \Bigl(u(X_i) - u(X_j)\Bigr)^2 K\biggl(\frac{\|x_i - x_j\|}{r}\biggr) + \frac{2L_K\wt{\delta}}{\wt{r}}\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \nonumber \\
+& \leq \biggl(1 + 4\bigl(2^{d + 2} - 1\bigr)\frac{\wt{\delta}}{r}\biggr) b_r(u) + \frac{2L_K\wt{\delta}}{\wt{r}}\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \nonumber \\
+& \leq \biggl(1 + 4\bigl(2^{d + 2} - 1\bigr)\frac{\wt{\delta}}{r}\biggr) b_r(u) + \frac{2L_K\wt{\delta}}{K(1)\wt{r}} \wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u\bigr) \nonumber 
+\end{align}
+or in other words
+\begin{align*}
+\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & \leq \biggl(1 - \frac{2L_K\wt{\delta}}{K(1)\wt{r}}\biggr)^{-1}\biggl(1 + 4\bigl(2^{d + 2} - 1\bigr)\frac{\wt{\delta}}{r}\biggr) b_r(u) \\
+& \leq \biggl(1 + \frac{2L_K}{K(1)} + 8\bigl(2^{d + 2} - 1\bigr)\biggr) \frac{\wt{\delta}}{r}b_r(u)
+\end{align*}
+where the second inequality follows from the algebraic identities $(1 - t)^{-1} \leq (1 + 2t)$ for any $0 < t < 1/2$ and $(1 + s)(1 + t) < 1 + 2s + t$ for any $0 < t < 1$ and $s > 0$.
+
+\subsection{Proof of Propositions~\ref{prop:dirichlet_energies} and~\ref{prop:isometry}}
+\label{subsec:proof_of_prop_dirichlet_energies_and_isometry}
+
+\paragraph{Proof of Proposition~\ref{prop:dirichlet_energies}.}
+Part (1) of Proposition~\ref{prop:dirichlet_energies} follows from
+\begin{align*}
+\sigma_K D_2(\Lambda_{r - 2\wt{\delta}} \wt{\mc{P}}^{\star}u) & \overset{(i)}{\leq} \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} E_{r - 2\wt{\delta}}(\wt{\mc{P}}^{\star}u) \\
+& \overset{(ii)}{\leq} \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \wt{E}_{r - 2\wt{\delta}}(\wt{\mc{P}}^{\star}u) \\
+& \overset{(iii)}{\leq} \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) b_r(u)
+\end{align*}
+where $(i)$ follows from Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, $(ii)$ follows from~\eqref{eqn:calder19_1}, and $(iii)$ follows from~Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}.
+
+Part (2) of Proposition~\ref{prop:dirichlet_energies} follows from
+\begin{align*}
+b_r(\wt{\mc{P}}f) & \overset{(iv)}{\leq} \wt{E}_{r + 2\wt{\delta}}(f) + \frac{2L_K}{K(1/2)}\frac{\wt{\delta}}{r}\wt{E}_{2(r + 2\wt{\delta})}(f)\\
+& \overset{(v)}{\leq} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl({E}_{r + 2\wt{\delta}}(f) + \frac{2L_K}{K(1/2)}\cdot\frac{\wt{\delta}}{r}{E}_{2(r + 2\wt{\delta})}(f)\Bigr) \\
+& \overset{(vi)}{\leq} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + L_p(2r + 4\wt{\delta})\Bigr) \cdot \Bigl(1 + \frac{2L_K}{K(1/2)}\frac{\wt{\delta}}{r}\Bigr) \cdot \sigma_{\eta} D_2(f)
+\end{align*}
+where $(iv)$ follows from Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}, $(v)$ follows from~\eqref{eqn:calder19_1}, and $(vi)$ follows from Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected}.
+The proposition follows upon taking $A_4 := 2L_K/K(1/2)$.
+
+\paragraph{Proof of Proposition~\ref{prop:isometry}.}
+\mbox{}\\
+\mbox{}\\
+\textit{Proof of (1).}
+We begin by upper bounding $\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}$. By the Cauchy-Schwarz inequality and the bound on $\|\wt{p}_n - p\|_{\infty}$ in~\eqref{eqn:calder19_1},
+\begin{align*}
+\Bigl|\wt{P}f(X_i)\Bigr|^2 & = n^2 \Bigl|\int_{U_i} f(x) \wt{p}_n(x) \,dx\Bigr|^2 \\
+& \leq n \int_{U_i} \bigl|f(x)\bigr|^2 \wt{p}_n(x) \,dx \\
+& \leq n \cdot \biggl[\int_{U_i} \bigl|f(x)\bigr|^2 p(x) \,dx + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \int_{U_i} \bigl|f(x)\bigr|^2 p(x) \,dx\biggr]
+\end{align*}
+and summing over $i = 1,\ldots,n$, we obtain
+\begin{equation}
+\label{pf:prop_isometry_1}
+\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}^2 \leq \biggl(1 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \bigl\|f\bigr\|_{\Leb^2(P)}^2.
+\end{equation}
+Now, noticing that $\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)} = \bigl\|\wt{P}^{\star}\wt{P}f\bigr\|_{\Leb^2(\wt{P}_n)}$, we can use the upper bound~\eqref{pf:prop_isometry_1} to show that
+\begin{align}
+\Bigl|\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}^2 - \bigl\|f\bigr\|_{\Leb^2(P)}^2\Bigr| & \leq \Bigl|\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}^2 - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2\Bigr| + \Bigl|\bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2 - \bigl\|f\bigr\|_{\Leb^2(P)}^2\Bigr| \nonumber \\
+& \overset{(i)}{\leq} \Bigl|\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}^2 - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2\Bigr|  + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{\Leb^2(P)}^2 \\
+& \overset{(ii)}{\leq} \biggl(2 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \Bigl|\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)} - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}\Bigr| \cdot \bigl\|f\bigr\|_{\Leb^2(P)} + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{\Leb^2(P)}^2 \nonumber \\
+& \leq \biggl(2 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)} \cdot \bigl\|f\bigr\|_{\Leb^2(P)} + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{\Leb^2(P)}^2 \label{pf:prop_isometry_2}
+\end{align}
+where $(i)$ follows from~\eqref{eqn:calder19_2} and $(ii)$ follows from~\eqref{pf:prop_isometry_1}.
+
+It remains to upper bound $\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)}^2$. Noting that $\wt{P}^{\star}\wt{P}f$ is piecewise constant over the cells $U_i$, we have
+\begin{equation*}
+\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)}^2 = \sum_{i = 1}^{n} \int_{U_i} \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx.
+\end{equation*}
+We will now establish the following Poincare-type inequality, 
+\begin{equation}
+\label{pf:prop_isometry_3}
+\int_{U_i} \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx \leq \frac{2^{d + 2}}{a_1(\Xset,d) \nu_d K(1/2)} \wt{\delta}^2 \wt{E}_{2\wt{\delta}}(f,U_i)
+\end{equation}
+which holds for all $i = 1,\ldots,n$. Summing up over $i$ implies
+\begin{equation*}
+\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)}^2 \leq \frac{2^{d+2}}{a_1(\Xset,d)K(1/2)p_{\min}} \cdot \wt{E}_{2\wt{\delta}}(f) \leq \frac{2^{d+3}}{a_1(\Xset,d)K(1/2)p_{\min}} A_1(\Xset,d,K,p_{\min}) D_2(f)
+\end{equation*}
+with the latter inequality following just as in the proof of Proposition~\ref{prop:dirichlet_energies}; the Lemma then follows by plugging this inequality into~\eqref{pf:prop_isometry_2}.
+
+Now it remains to show~\eqref{pf:prop_isometry_3}. The key point is that for any pair of points $x,x' \in \Xset$ such that $\|x - x'\| \leq r$  the triangle inequality
+\begin{equation*}
+\bigl|f(x') - f(x)\bigr|^2 \leq 2\bigl|f(x') - f(z)\bigr|^2 + 2\bigl|f(z) - f(x)\bigr|^2
+\end{equation*}
+holds for each $z \in B\bigl(x,\wt{\delta}\bigr) \cap B\bigl(x',\wt{\delta}\bigr) \cap \Xset =: W$, and we have that
+\begin{align*}
+\bigl|f(x') - f(x)\bigr|^2 & \leq \frac{2}{\vol(W)}\biggl[\int_{W} \bigl|f(x') - f(z)\bigr|^2 \,dz + \int_{W} \bigl|f(z) - f(x)\bigr|^2 \,dz\biggr] \\
+& \leq \frac{2}{\vol(W)K(1/2)}\biggl[\int_{\Xset} K\biggl(\frac{x' - z}{2\wt{\delta}}\biggr)\bigl|f(x') - f(z)\bigr|^2 \,dz + \int_{\Xset} K\biggl(\frac{z - x}{2\wt{\delta}}\biggr)\bigl|f(z) - f(x)\bigr|^2 \,dz\biggr] \\
+& \leq \frac{2^{d + 1}}{a_1(\Xset,d) \nu_d \wt{\delta}^d K(1/2)}\biggl[\int_{\Xset} K\biggl(\frac{x' - z}{2\wt{\delta}}\biggr)\bigl|f(x') - f(z)\bigr|^2 \,dz + \int_{\Xset} K\biggl(\frac{z - x}{2\wt{\delta}}\biggr)\bigl|f(z) - f(x)\bigr|^2 \,dz\biggr]
+\end{align*}
+where the last inequality follows from Assumption~\ref{asmp:r_small_1}. We use this along with the Cauchy-Schwarz inequality to prove~\eqref{pf:prop_isometry_3}:
+\begin{align*}
+\int_{U_i} & \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx \\
+& =\int_{U_i} \biggl(n\cdot\int_{U_i} \bigl(f(x) - f(x')\bigr) \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx \\
+& \leq n \cdot \int_{U_i} \int_{U_i} \bigl(f(x) - f(x')\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx \\
+& \leq \frac{2^{d + 2}}{a_1(\Xset,d) \nu_d \wt{\delta}^d K(1/2)}n\int_{U_i} \int_{U_i} \int_{\Xset} K\biggl(\frac{x - z}{2\wt{\delta}}\biggr)\bigl|f(x) - f(z)\bigr|^2 \wt{\rho}_n(x') \wt{\rho}_n(x)  \,dx' \,dz \,dx \\
+& \leq \frac{2^{d + 2}}{a_1(\Xset,d) \nu_d \wt{\delta}^d K(1/2) p_{\min}} \int_{U_i} \int_{U_i} K\biggl(\frac{x - z}{2\wt{\delta}}\biggr)\bigl|f(x) - f(z)\bigr|^2 \wt{\rho}_n(z) \wt{\rho}_n(x)  \,dz \,dx \\
+& =  \frac{2^{d + 2}}{a_1(\Xset,d) \nu_d K(1/2) p_{\min}} \wt{\delta}^2 \wt{E}_{2\wt{\delta}}(f,U_i).
+\end{align*}
+
+\textit{Proof of (2).}
+By the triangle inequality,
+\begin{align}
+\Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(P)}^2 - \|u\|_{\Leb^2(P_n)}^2\Bigr| & \leq \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(P)}^2 - \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2\Bigr| + \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 - \|u\|_{\Leb^2(P_n)}^2\Bigr| \nonumber \\
+& \leq A_1(\theta + \wt{\delta}) \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 + \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 - \|u\|_{\Leb^2(P_n)}^2\Bigr| \nonumber\\
+& = A_1(\theta + \wt{\delta}) \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 + \Bigl(\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} + \|u\|_{\Leb^2(P_n)}\Bigr) \cdot \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|u\|_{\Leb^2(P_n)}\Bigr| \label{pf:prop_isometry_4}
+\end{align}
+To upper bound the second term in the above expression, we first note that~$\|u\|_{\Leb^2(P_n)} = \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}$, and thus
+\begin{align}
+\Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|u\|_{\Leb^2(P_n)} \Bigr| & = \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}\Bigr| \nonumber \\
+& \overset{(iii)}{\leq} \|\Lambda_{\wt{r}}\wt{\mc{P}}^{\star}u - \wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)} \nonumber \\
+& \overset{(iv)}{\leq} r \sqrt{\frac{2}{\sigma_K p_{\min}} E_{\wt{r}}(\wt{\mc{P}}^{\star}u)} \nonumber \\
+& \overset{(v)}{\leq} r \sqrt{\frac{2}{\sigma_K p_{\min}} \Bigl(1 + A_3\frac{\wt{\delta}}{r}\Bigr) b_r(u)} \label{pf:prop_isometry_5}
+\end{align}
+where $(iii)$ follows by the triangle inequality, $(iv)$ follows from Lemma~\ref{lem:smoothening_error}, and $(v)$ follows from~\eqref{eqn:calder19_1} and Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}.
+On the other hand, by~\eqref{eqn:calder19_2} and Lemma~\ref{lem:smoothening_error},
+\begin{align*}
+\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 & \leq \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \|\wt{\mc{I}}u\|_{\Leb^2(P)}^2 \\
+& \leq \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(P)}^2 \\
+& \leq \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)^2 \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}^2 \\
+& = \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)^2 \|u\|_{\Leb^2(P_n)}^2
+\end{align*}
+and plugging this estimate along with~\eqref{pf:prop_isometry_5} back into~\eqref{pf:prop_isometry_4}, we obtain part (2) of Proposition~\ref{prop:isometry}.
+
+\subsection{\textcolor{red}{(TODO)}: Move these.}
+
+In the above we used the following:
+\begin{enumerate}[label=(OT\arabic*)]
+	\item
+	\label{asmp:r_small_1} For all $x,x' \in \Xset$ such that $\|x - x'\| \leq r$, we have that for $W = B(x,r) \cap B(x',r) \cap \Xset$,
+	\begin{equation*}
+	\vol(W) \geq a_1(\Xset,d) \cdot \nu_d \biggl(\frac{r}{2}\biggr)^d
+	\end{equation*}
+	for some constant $a_1(\Xset,d)$. 
+	\item 
+	\label{asmp:r_small_3} For any $x \in \Xset$, letting $\mc{B}_x:= (\Xset - x)/r~ \cap~B(0,1)$, we have that $\int_{\mc{B}_x} \psi(\|z\|) \,dz \geq a_1(\Xset,d) \cdot \int_{B(0,1)} \psi(\|z\|) \,dz$.
+\end{enumerate}
 
@@
 
+We use Lemma~\ref{lem:empirical_norm_sobolev} to lower bound $\norm{f_0}_n^2$ by (a constant times) the $\Leb^2(\Xset)$ norm of $f$.
+
+\begin{lemma}
+	\label{lem:empirical_norm_sobolev}
+	Suppose $P$ satisfies~\ref{asmp:bounded_lipschitz_density}. If $f \in H^1(\Xset,M)$ is lower bounded in $\Leb^2(\Xset)$ norm,
+	\begin{equation}
+	\label{eqn:empirical_norm_sobolev_1}
+	\norm{f}_{\Leb^2(\Xset)} \geq 
+	\begin{cases*}
+	\frac{C_6 M}{\delta} \cdot \max\Bigl\{n^{-1/2},n^{-1/d}\Bigr\},& ~~\textrm{if $2 \neq d$} \\
+	\frac{C_6 M}{\delta} \cdot n^{-a/2},& ~~\textrm{if $2 = d$, for any $0 < a < 1$}
+	\end{cases*}
+	\end{equation}
+	for some $\delta \leq 1$, then
+	\begin{equation}
+	\label{eqn:empirical_norm_sobolev}
+	\norm{f}_n^2 \geq \delta \cdot \Ebb\Bigl[\norm{f}_n^2\Bigr] 
+	\end{equation}
+	with probability at least $1 - 5 \delta$.
+\end{lemma}
+
+\paragraph{Proof of Lemma~\ref{lem:empirical_norm_sobolev}.}
+To prove~\eqref{eqn:empirical_norm_sobolev} we will show
+\begin{equation*}
+\mathbb{E}\bigl[\norm{f}_n^4\bigr] \leq \left(1 + \frac{1}{b^2}\right)\cdot \left(\mathbb{E}\bigl[\norm{f}_n^2\bigr]\right)^2
+\end{equation*}
+whence the claim follows from the Paley-Zygmund inequality (Lemma~\ref{lem:paley_zygmund}).
+Since $p \leq p_{\max}$ is uniformly bounded, we can relate $\mathbb{E}\bigl[\norm{f}_n^4\bigr]$ to the $\Leb^4$ norm,
+\begin{equation*}
+\mathbb{E}\bigl[\norm{f}_n^4\bigr] = \frac{(n-1)}{n}\left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + \frac{\mathbb{E}\Bigl[\bigl(f(x_1)\bigr)^4\Bigr]}{n} \leq \left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + p_{\max}^2\frac{\norm{f}_{\Leb^4}^4}{n}.
+\end{equation*}
+We will use a Sobolev inequality to relate $\norm{f}_{\Leb^4}$ to $\norm{f}_{H^1(\Xset)}$.
+The nature of this inequality depends on the relationship between $s$ and $d$ (see Theorem 6 in Section 5.6.3 of \cite{evans10} for a formal statement), so from this point on we divide our analysis into three cases: (i) the case where $2s > d$, (ii) the case where $2s < d$, and (iii) the borderline case $2s = d$.
+
+\textit{Case 1: $2s > d$.}
+When $2s > d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in C^{\gamma}(\overline{\Xset})$ for some $\gamma > 0$ which depends on $s$ and $d$, with the accompanying estimate
+\begin{equation*}
+\sup_{x \in \Xset} \abs{f(x)} \leq \norm{f}_{C^{\gamma}(\Xset)} \leq c \norm{f}_{H^s(\Xset)}.
+\end{equation*}
+Therefore,
+\begin{align*}
+\norm{f}_{\Leb^4}^4 & = \int_{\Xset} [f(x)]^4 \,dx \\
+& \leq \left(\sup_{x \in \Xset} \abs{f(x)}\right)^2 \cdot \int_{\Xset} [f(x)]^2 \,dx \\
+& \leq c \norm{f}_{H^s(\Xset)}^2 \cdot \norm{f}_{\Leb^2(\Xset)}^2.
+\end{align*}
+Since by assumption
+\begin{equation*}
+\norm{f}_{\Leb^2(\Xset)}^2 \geq c_1^2 \cdot b^2 \cdot \norm{f}_{W_d^{s,2}(\Xset)}^2 \cdot \frac{1}{n},
+\end{equation*}
+we have
+\begin{equation*}
+p_{\max}^2\frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c \norm{f}_{H^s(\Xset)}^2 \cdot \frac{\norm{f}_{\Leb^2(\Xset)}^4}{n \norm{f}_{\Leb^2(\Xset)}^2} \leq c \frac{\norm{f}_{\Leb^2(\Xset)}^4}{c_1^2 b^2} \leq \frac{\Ebb\bigl[\norm{f}_n^2\bigr]}{b^2},
+\end{equation*}
+where the last inequality follows by taking $c_1$ sufficiently large.
+
+\textit{Case 2: $2s < d$.}
+When $2s < d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in \Leb^q(\Xset)$ for $q = 2d/(d - 2s)$, and moreover that
+\begin{equation*}
+\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{H^s(\Xset)}.
+\end{equation*}
+Since $4 = 2\theta + (1 - \theta)q$ for $\theta = 2 - d/(2s)$, Lyapunov's inequality implies
+\begin{equation*}
+\norm{f}_{\Leb^4(\Xset)}^4 \leq \norm{f}_{\Leb^2}^{2\theta} \cdot \norm{f}_{\Leb^q(\Xset)}^{(1 - \theta)q} \leq c \norm{f}_{\Leb^2(\Xset)}^{4} \cdot \left(\frac{\norm{f}_{H^s(\Xset)}}{\norm{f}_{\Leb^2(\Xset)}}\right)^{d/s}.
+\end{equation*}
+By assumption, $\norm{f}_{\Leb^2(\Xset)} \geq c_1 b \norm{f}_{H^s(\Xset)} n^{-s/d}$, and therefore
+\begin{equation*}
+p_{\max}^2 \frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c\norm{f}_{\Leb^2(\Xset)}^4 \left(\frac{\norm{f}_{H^s(\Xset)}}{n^{s/d}\norm{f}_{\Leb^2(\Xset)}}\right)^{d/s} \leq \frac{c\norm{f}_{\Leb^2(\Xset)}^4}{c_1b^{d/s}} \leq \frac{\norm{f}_{\Leb^2(\Xset)}^4}{b^2}.
+\end{equation*}
+where the last inequality follows when $c_1$ is sufficiently large, and keeping in mind that $d/s > 2$ and $b \geq 1$.
+
+\textit{Case 3: $2s = d$.}
+Assume $f$ satisfies~\eqref{eqn:paley_zygmund_1} for a given $0 < a < 1$.
+When $2s = d$, since $\Xset$ is a Lipschitz domain we have that $f \in L^q(\Xset)$ for any $q < \infty$, with the accompanying estimate
+\begin{equation*}
+\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{H^s(\Xset)}.
+\end{equation*}
+In particular the above holds for $q = 2/(1 - a)$ when $1/2 < a < 1$, and for any $q > 4$ when $0 < a < 1/2$. Using Lyapunov's inequality as in the previous case then implies the desired result.
+
+\paragraph{Proof of Lemma~\ref{lem:empirical_norm_sobolev}.}
+
+\textcolor{red}{(TODO)}
+
 \section{Proof of theorems}
+
+\subsection{Proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}}
+\label{subsec:laplacian_smoothing_estimation1_pf}
+Let $\mc{E}$ be the set of $\mathbf{X} = (X_1,\ldots,X_n)$ for which
+\begin{enumerate}[(i)]
+	\item
+	\label{pf:laplacian_smoothing_estimation1_0}
+	$f_0^T \Lap_{n,r} f_0 \leq \frac{p_{\max}^2 \sigma_K}{\delta} n^2 r^{d + 2}M^2$
+	\item 
+	\label{pf:laplacian_smoothing_estimation1_1}
+	$c_3 \cdot \min\{k^{2/d}nr^{d+2},nr^d\} \leq \lambda_k(G_{n,r}) \leq C_3k^{2/d}nr^{d + 2} $ for all $2 \leq k \leq n$.
+\end{enumerate}
+We already know, from Lemma~\ref{lem:graph_sobolev_seminorm} and Corollary~\ref{cor:neighborhood_eigenvalue}, that for any $\delta > 0$ and $(\log(n)/n)^{1/d} \leq r \leq a_0$,
+\begin{equation}
+\label{pf:laplacian_smoothing_estimation1_1.5}
+\Pbb\bigl({\bf X} \in \mc{E}\bigr) \geq 1 - \delta - A_1n\exp\bigl(-a_1nr^d\bigr).
+\end{equation}
+We will now show that conditional on $\mathbf{X} \in \mc{E}$, the empirical mean square error is small with high probability.
+Having conditioned on $\mathbf{X}$, we can use Lemma~\ref{lem:ls_fixed_graph_estimation} to control the remaining randomness arising from the noisy response $Y$, which in this context implies
+\begin{equation}
+\label{pf:laplacian_smoothing_estimation1_2}
+\begin{aligned}
+\Pbb\Biggl(& \frac{1}{n}\bigl\|\wt{f}_{\LS}(G_{n,r}) - f_0\bigr\|_2^2 \leq \frac{2\rho}{n} \bigl(f_0^T \Lap_{n,r} f_0\bigr) + \frac{10}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k}(G_{n,r}) + 1\bigr)^2} \Bigg| \mathbf{X} = x \in \mc{E}\Biggr) \\
+& \geq 1 - \exp\Biggl\{-\sum_{k = 1}^{n}\frac{1}{\bigl(\rho \lambda_k(G_{n,r}) + 1\bigr)^2}\Biggr\} \\
+& \geq 1 - \exp\Biggl\{-\sum_{k = 1}^{n}\frac{1}{\bigl(C_3M^{-4/(2+d)}k^{2/d}n^{-2/(2+d)} + 1\bigr)^2}\Biggr\} \\
+& \geq 1 - \exp\Bigl\{-c M^{d/(2d + 4)} n^{d/(2+d)}\Bigr\}
+\end{aligned}
+\end{equation}
+where the second inequality follows from the upper bound in~\ref{pf:laplacian_smoothing_estimation1_1} and the choice of regularization parameter $\rho = M^{-4/(2+d)}(nr^{d+2})^{-1}n^{-2/(2+d)}$.
+
+It remains only to upper bound the (conditional on ${\bf X} = x$ for some $x \in \mc{E}$) squared bias and variance terms.
+A sufficient upper bound on the bias term comes immediately from the upper bound~\ref{pf:laplacian_smoothing_estimation1_0} and our choice of $\rho$,
+\begin{equation}
+\label{pf:laplacian_smoothing_estimation1_2.5}
+\frac{\rho}{n} \bigl(f_0^T \Lap_{{n,r}} f_0\bigr) \leq \frac{p_{\max}^2 \sigma_K}{\delta} M^{2/(2+d)} n^{-2/(2 + d)}.
+\end{equation}
+
+To upper bound the variance term, we replace the eigenvalues $\lambda_k(G_{n,r})$ by their lower bounds in~\ref{pf:laplacian_smoothing_estimation1_1} and plug in our choice of $\rho$ to obtain,
+\begin{align}
+\frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k}(G_{n,r}) + 1\bigr)^2} & \leq \frac{C}{n}\sum_{k = 1}^{n} \biggl\{\frac{1}{(\rho k^{2/d} n r^{d + 2} + 1)^2} + \frac{1}{(\rho n r^{d} + 1)^2} \biggr\} \nonumber \\
+& \leq \frac{C}{n}\sum_{k = 1}^{n} \biggl\{\frac{1}{\bigl(M^{-4/(2 + d)}n^{-2/(2+d)}k^{2/d} + 1\bigr)^2}\biggr\} + Cn^{4/(2 + d)}M^{8/(2+d)}r^{4} \nonumber \\
+& \leq CM^{2d/(2 + d)} n^{-2/(2 + d)} + CM^{8/(2 + d)}n^{(2 - d)/(2 + d)} \sum_{k = k_{\star}}^{n} \Bigl\{\frac{1}{k^{4/d}}\Bigr\}  + CM^{8/(2 + d)}n^{4/(2 + d)}r^4 \label{pf:laplacian_smoothing_estimation1_2.25}
+\end{align}
+where $k_{\star} = M^{2d/(2 + d)}n^{d/(2 + d)}$.
+
+We now upper bound the 2nd and 3rd term on the right hand side of~\eqref{pf:laplacian_smoothing_estimation1_2.25}.
+To control the 3rd term, we use the upper bound $r \leq n^{-3/(4 + 2d)} M^{(d - 4)/(4 + 2d)}$ assumed in~\ref{asmp:ls_kernel_radius_estimation} to determine that
+\begin{equation}
+\label{pf:laplacian_smoothing_estimation1_3}
+M^{8/(2 + d)}n^{4/(2 + d)}r^4 \leq M^{2d/(2 + d)}n^{-2/(2 + d)} 
+\end{equation}
+To control the 2nd term on the right hand side of~\eqref{pf:laplacian_smoothing_estimation1_2.25}, we treat it as a Riemann sum evaluated at the right end points of $[k_{\star},k_{\star} + 1], [n - 1,n]$.
+We use the fact that such a Riemann sum upper bounds the integral of a montonically nonincreasing function to conclude that
+\begin{equation}
+\label{pf:laplacian_smoothing_estimation1_4}
+\begin{aligned}
+\sum_{k = k_{\star}}^{n} \frac{1}{k^{4/d}} & \leq \int_{k_{\star}}^{n} \frac{1}{x^{4/d}} \,dx \\
+& = - \frac{1}{(4/d - 1)} x^{-(4/d - 1)} \Big|_{k_{\star}}^{n} \\
+& \leq \frac{1}{(4/d - 1)} k_{\star}^{-(4/d - 1)}
+\end{aligned}
+\end{equation}
+where the equality in the above computation holds because $d < 4$.
+Recalling the definition of $k_{\star}$, plugging this back into~\eqref{pf:laplacian_smoothing_estimation1_3} implies a sufficient upper bound on the variance term.
+Along with~\eqref{pf:laplacian_smoothing_estimation1_1.5}-\eqref{pf:laplacian_smoothing_estimation1_2.5}, this establishes the claim.
+
+\paragraph{Proving near-optimal bounds when $d = 4$.}
+Note that nothing in the derivations of  ~\eqref{pf:laplacian_smoothing_estimation1_1.5}-\eqref{pf:laplacian_smoothing_estimation1_2.5} relied on $d < 4$, and so they all hold when $d = 4$.
+We now upper bound the 2nd and 3rd terms in~\eqref{pf:laplacian_smoothing_estimation1_2.5}.
+By setting $r = C_4(\log(n)/n)^{1/4}$ for some constant $C_4$, we have that
+\begin{equation*}
+M^{8/(2+d)}n^{4/(2+d)}r^4 = C_4^4 M^{8/(2+d)} \log(n) n^{-1/3},
+\end{equation*}
+taking care of the third term. On the other hand, we have that
+\begin{equation*}
+\sum_{k = k_{\star}}^{n} \frac{1}{k} \leq \log(n)
+\end{equation*}
+and so the second term is also upper bounded by $C M^{8/(2+d)} \log(n) n^{-1/3}$.
+
+\subsection{Proof of Theorem~\ref{thm:laplacian_smoothing_testing}}
+In this proof we will need to invoke Lemma~\ref{lem:empirical_norm_sobolev}, and it is in order to satisfy the conditions of this Lemma that we require $M \leq n^{(4 - d)/(4 + d)}$.
+More specifically, by~\eqref{eqn:laplacian_smoothing_testing} along with the restriction $M \leq n^{(4 - d)/(4 + d)}$,
+\begin{equation*}
+\norm{f}_{\Leb^2(\Xset)} \geq \frac{C}{\delta^2} M^{2d/(4 + d)} n^{-4/(4 + d)} \geq \frac{C}{\delta^2} M^2 n^{-2/d}
+\end{equation*}
+and~\eqref{eqn:empirical_norm_sobolev_1} is therefore satisfied when we choose $C \geq C_6$ in~\eqref{eqn:laplacian_smoothing_testing}.
+
+We proceed to prove Theorem~\ref{thm:laplacian_smoothing_testing}.
+Throughout this proof we set $\delta = 1/b$.
+We will proceed, as in Subsection~\ref{subsec:laplacian_smoothing_estimation1_pf}, by conditioning on the random design $\mathbf{X}$.
+Let $\mc{E}(f_0)$ be the set of $\mathbf{X} \in \Xset^n$ such that~\eqref{eqn:graph_sobolev_seminorm},~\eqref{eqn:neighborhood_eigenvalue_2}, and~\eqref{eqn:empirical_norm_sobolev_1} hold: by Lemmas~\ref{lem:graph_sobolev_seminorm} and~\ref{lem:empirical_norm_sobolev}, and Corollary \ref{cor:neighborhood_eigenvalue},
+\begin{equation*}
+\Pbb\bigl(\mc{E}(f_0)\bigr) \geq 1 - 6\delta - A_1n\exp\bigl\{-a_1nr^d\bigr\}. 
+\end{equation*}
+For any $\mathbf{X} \in \mc{E}(f_0)$ the conditional variance of $T_{\LS}$ can be upper bounded using~\eqref{eqn:neighborhood_eigenvalue_2} and~\ref{asmp:ls_kernel_radius_testing},
+\begin{align*}
+\frac{1}{n} \biggl(\sum_{k = 1}^{n} \bigl[\lambda_k(G_{n,r}) + 1\bigr]^{-4}\biggr)^{1/2} & \leq \frac{1}{c_5^2 n} \biggl(M^{\frac{32}{8 + d}}r^8n^{\frac{20+d}{4+d}} + \sum_{k = 1}^{n} \bigl[n^{-\frac{4}{4 + d}}k^{\frac{2}{d}}M^{-\frac{8}{4 + d}} + 1\bigr]^{-4}\biggr)^{1/2} \\
+& \leq \frac{1}{c_5^2 n} \biggl(M^{\frac{4d}{4 + d}}n^{\frac{2d}{4 + d}} + \sum_{k = 1}^{n} \bigl[n^{-\frac{4}{4 + d}}k^{\frac{2}{d}}M^{-\frac{8}{4 + d}} + 1\bigr]^{-4}\biggr)^{1/2} \\
+& \leq \frac{1}{c_5^2 n} \biggl(3 M^{\frac{4d}{4 + d}}n^{2d/(4 + d)}\biggr)^{1/2} \\
+& = \frac{\sqrt{3}}{c_5^2} M^{\frac{2d}{4 + d}} n^{-\frac{4}{4 + d}}.
+\end{align*}
+The last inequality in the above display similarly to~\eqref{pf:laplacian_smoothing_estimation1_4}. Setting $k_{\star} = n^{2d/(4+d)}M^{4d/(4 + d)}$, we derive that
+\begin{equation}
+\begin{aligned}
+\label{pf:laplacian_smooting_testing1}
+\sum_{k = 1}^{n} \bigl[n^{-\frac{4}{4 + d}}k^{\frac{2}{d}}M^{-\frac{8}{4 + d}} + 1\bigr]^{-4} & \leq k_{\star} + n^{\frac{16}{4 + d}} M^{\frac{32}{4 + d}}\sum_{k = k_{\star}}^{n} k^{-\frac{8}{d}}\\
+& \leq k_{\star} + n^{\frac{16}{4 + d}} M^{\frac{32}{4 + d}} \int_{k_{\star}}^{n} x^{-\frac{8}{d}} \,dx \\
+& \leq  k_{\star} + \frac{1}{(8/d - 1)}n^{\frac{16}{4 + d}} M^{\frac{32}{4 + d}} k_{\star}^{1 - 8/d} \\
+& \leq 2k_{\star}.
+\end{aligned}
+\end{equation}
+As a result, 
+\begin{align*}
+\frac{2\rho}{n} \bigl(f_0^T \Lap_{n,r} f_0) + \frac{4}{\delta n}\biggl(\sum_{k = 1}^{n} \bigl[\lambda_k(G_{n,r}) + 1\bigr]^{-4}\biggr)^{1/2} & \leq \frac{1}{\delta}\biggl(2C_1 + \frac{4\sqrt{3}}{c_5^2}\biggr)n^{-\frac{4}{4 + d}} M^{\frac{2d}{4 + d}} \\
+& \leq \delta p_{\min} \norm{f}_{\Leb^2(\Xset)}^2 \leq \delta \Ebb\bigl[\norm{f}_n^2\bigr] \\
+& \leq \norm{f}_n^2,
+\end{align*}
+where the second inequality follows upon choosing $C \geq p_{\min}^{-1}(2C_1 + 4\sqrt{3}/c_5^2)$ in~\eqref{eqn:laplacian_smoothing_testing}.
+We may therefore bound the Type II error of $\wt{\phi}_{\LS}$ by appealing to Lemma~\ref{lem:ls_fixed_graph_testing},
+\begin{align*}
+\Ebb_{f_0}\Bigl[1 - \wt{\phi}_{\LS}\Bigr] & \leq \Ebb_{f_0}\Bigl[1 - \wt{\phi}_{\LS}\big| \mathbf{X} \in \mc{E}(f_0)\Bigr] + 1 - \Pbb\Bigl(\mc{E}(f_0)\Bigr) \\
+& \leq 4 \delta^2 + 8 \delta \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G_{n,r}) + 1)^4} \Biggr)^{-1/2} + 6\delta + A_1\exp\bigl\{-a_1nr^{d}\bigr\} \\
+& \leq 4 \delta^2 + 16\delta n^{-d(4 + d)}M^{-2d/(4 + d)}  + 6\delta + A_1\exp\bigl\{-a_1nr^{d}\bigr\}
+\end{align*}
+where the last inequality follows from the reversing the inequalities in~\eqref{pf:laplacian_smooting_testing1}, with a factor of $1/2$ instead of $2$ in front.
+
+\subsection{Proof of~\eqref{eqn:laplacian_smoothing_testing_low_smoothness}}
+When $\rho = 0$, the Laplacian smoother $\wh{f}_{\LS} = Y$, the test statistic $T_{\LS} = \frac{1}{n}\|Y\|_2^2$, and the threshold $\wh{t}_b = 1 + 2bn^{-1/2}$. The expectation of $T_{\LS}$ is 
+\begin{equation*}
+\Ebb\bigl[T_{\LS}\bigr] = \mathbb{E}\bigl[f_0^2(X)\bigr] + 1 \geq p_{\min} \norm{f_0}_{\Leb^2(\Xset)}^2 + 1
+\end{equation*}
+When $f_0 \in \Leb^4(\Xset,M)$, the variance can be upper bounded
+\begin{equation*}
+\Var\bigl[T_{\LS}\bigr] \leq \frac{1}{n}\Bigl(3 + p_{\max} M^4 + p_{\max}\norm{f_0}_{\Leb^2(\Xset)}^2\Bigr).
+\end{equation*}
+Now, let us assume that
+\begin{equation*}
+\norm{f_0}_{\Leb^2(X)}^2 \geq \frac{4b}{p_{\min}} n^{-1/2},
+\end{equation*}
+so that $E[T_{\LS}] - \wh{t}_b \geq E[T_{\LS}]/2$. Hence, by Chebyshev's inequality
+\begin{align*}
+\mathbb{P}_{f_0}\Bigl(T_{\LS} \leq \wh{t}_b \Bigr) & \leq 4 \frac{\Var_{f_0}\bigl[T_{\LS}\bigr]}{\mathbb{E}[f^2(X)]^2} \\
+& \leq \frac{4}{n} \cdot \frac{ 3 + p_{\max}\bigl(M^4 + \|f_0\|_{\Leb^2(\Xset)}^2 \bigr)}{p_{\min}^2 \|f_0\|_{\Leb^2(\Xset)}^4} \\
+& \leq \frac{1}{16b^2}\Bigl(3 + p_{\max}n^{-1/2} + p_{\max}M^4\Bigr).
+\end{align*}
+
+\section{Concentration Inequalities}
+\begin{lemma}
+	\label{lem:chi_square_bound}
+	Let $\xi_1,\ldots,\xi_N$ be independent $N(0,1)$ random variables, and let $U := \sum_{k = 1}^{N} a_k(\xi_k^2 - 1)$.  Then for any $t > 0$,
+	\begin{equation*}
+	\Pbb\Bigl[U \geq 2 \norm{a}_2 \sqrt{t} + 2 \norm{a}_{\infty}t\Bigr] \leq \exp(-t).
+	\end{equation*}
+	In particular if $a_k = 1$ for each $k = 1,\ldots,N$, then
+	\begin{equation*}
+	\Pbb\Bigl[U - N \geq 2\sqrt{N t} + 2t\Bigr] \leq \exp(-t).
+	\end{equation*}
+\end{lemma}
+
+The proof of Lemma~\ref{lem:empirical_norm_sobolev} relies on (a variant of) the Paley-Zygmund Inequality.
+\begin{lemma}
+	\label{lem:paley_zygmund}
+	Let $f$ satisfy the following moment inequality for some $b \geq 1$:
+	\begin{equation}
+	\label{eqn:paley_zygmund_1}
+	\Ebb\bigl[\norm{f}_n^4\bigr] \leq \left(1 + \frac{1}{b^2}\right)\cdot\Bigl(\Ebb\bigl[\norm{f}_n^2\bigr]\Bigr)^2.
+	\end{equation}
+	Then,
+	\begin{equation}
+	\label{eqn:paley_zygmund_2}
+	\mathbb{P}\left[\norm{f}_n^2 \geq \frac{1}{b} \Ebb\bigl[\norm{f}_n^2\bigr]\right] \geq 1 - \frac{5}{b}.
+	\end{equation}
+\end{lemma}
+\begin{proof}
+	Let $Z$ be a non-negative random variable such that $\mathbb{E}(Z^q) < \infty$. The Paley-Zygmund inequality says that for all $0 \leq \lambda \leq 1$,
+	\begin{equation}
+	\label{eqn:paley_zygmund_pf1}
+	\mathbb{P}(Z > \lambda \mathbb{E}(Z^p)) \geq \left[(1 - \lambda^p) \frac{\mathbb{E}(Z^p)}{(\mathbb{E}(Z^q))^{p/q}}\right]^{\frac{q}{q - p}}
+	\end{equation}
+	Applying~\eqref{eqn:paley_zygmund_pf1} with $Z = \norm{f}_n^2$, $p = 1$, $q = 2$ and $\lambda = \frac{1}{b}$, by assumption~\eqref{eqn:paley_zygmund_1} we have
+	\begin{equation*}
+	\mathbb{P}\Bigl(\norm{f}_n^2 > \frac{1}{b} \mathbb{E}[\norm{f}_n^2]\Bigr) \geq \Bigl(1 - \frac{1}{b}\Bigr)^2 \cdot  \frac{\bigl(\mathbb{E}[\norm{f}_n^2]\bigr)^2}{\mathbb{E}[\norm{f}_n^4]} \geq \frac{\Bigl(1 - \frac{2}{b}\Bigr)}{\Bigl(1 + \frac{1}{b^2}\Bigr)} \geq 1 - \frac{5}{b}.
+	\end{equation*}
+\end{proof}
+
+\clearpage
 

### 2020-10-06 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing2.tex
Fixed a few typos, and shortened one wandering paragraph.

@@
 \title{Minimax optimal Laplacian smoothing}
-\author{Alden Green, Sivaraman Balakrishnan, and Ryan Tibshirani}
+\author{Alden Green, Sivaraman Balakrishnan, and Ryan J. Tibshirani}
 \date{\today}
@@
 \end{itemize}
+\textcolor{red}{(TODO)}: Perhaps additional motivation is needed?
 
@@
 
-Works in the first category view the responses $Y = \theta + \varepsilon$ as noisy samples of a regression function $\theta \in \Reals^n$ defined on the vertices of \emph{fixed graph} $G$.
+Works in the first category view the responses $Y = \theta + \varepsilon$ as noisy samples of a regression function $\theta \in \Reals^n$ defined on the vertices of a \emph{fixed graph} $G$.
 and bound the statistical error as a function of $\theta$ and $G$.
@@
 \end{equation}
-as $n \to \infty$ and $r \to 0$. This is meant in various senses:
-\begin{itemize}
-	\item Pointwise convergence, meaning that $\Lap_{n,r}f \to \Delta_Pf$ in an appropriate norm \citep{belkin03,belkin05,lafon04,hein05,singer06,gine06}. 
-	\item Spectral convergence, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$, where the latter are the Dirichlet eigenvalues of $\Delta_P$ \citep{belkin07,burago2014,trillos2018,trillos2019,calder2019}.\footnote{\label{footnote:dirichlet_eigenvalues} Formally, Dirichlet eigenvalues are solutions to the problem
+as $n \to \infty$ and $r \to 0$.
+This is meant in (at least) three senses: (1) \emph{pointwise convergence}, meaning that $\Lap_{n,r}f \to \Delta_Pf$ in an appropriate norm \citep{belkin03,belkin05,lafon04,hein05,singer06,gine06}; (2) spectral convergence, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$, where the latter are the Dirichlet eigenvalues of $\Delta_P$ \citep{belkin07,burago2014,trillos2018,trillos2019,calder2019}.\footnote{\label{footnote:dirichlet_eigenvalues} Formally, Dirichlet eigenvalues are solutions to the problem
 	$$
@@
 	\restr{g}{\partial \Xset} = 0
-		$$} 
-	\item Convergence of norms, meaning $f^T \Lap_{n,r} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$ \citep{bousquet03,hein06,zhou11}. 
-\end{itemize}
+	$$}, and (3) convergence of norms, meaning $f^T \Lap_{n,r} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$ \citep{bousquet03,hein06,zhou11}.
 While these results imply \emph{consistency} of many different graph Laplacian based procedures, the statistical \emph{optimality} of Laplacian smoothing over neighborhood graphs is still not well understood.
@@
 	\item In Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we leverage the results of Section~\ref{sec:graph_sobolev_classes} to derive tight upper bounds on the error of Laplacian smoothing methods.
-	\item In Section~\ref{sec:simulations} we provide some experiments supporting our theory, and conclude in Section~\ref{sec:discussion} with some potential methodological and theoretical extensions.
+	\item In Section~\ref{sec:simulations} we provide some experiments supporting our theory, before concluding in Section~\ref{sec:discussion}.
 \end{itemize}
@@
 We use $A,A_1, \ldots$ and $C,C_1, \ldots$ to refer to constants which do not depend on $n$, $r$, or $f_0$; likewise with $a, a_1,\ldots$ and $c, c_1,\ldots$.
+
+\textcolor{red}{(TODO)}: Complete this.
 
@@
 
-Before we get to our main results, we briefly review minimax optimal estimation and testing rates over Sobolev classes.
+Before we get to our main results, we review minimax optimal estimation and testing rates over Sobolev classes.
 We also single out a notable method which achieves these rates: \emph{smoothing splines}.
@@
 
-When $d = 1$, the embedding of $H^1(\Xset)$ into $C^{0,q}(\Xset)$ facilitates a general analysis of (penalized) least estimators over $H^1(\Xset)$, using arguments from empirical process theory.
+% AJG commented this out.
+% When $d = 1$, the embedding of $H^1(\Xset)$ into $C^{0,q}(\Xset)$ facilitates a general analysis of (penalized) least estimators over $H^1(\Xset)$, using arguments from empirical process theory.
 However when $d \geq 2$, such arguments fall apart: for example, good bounds on the metric entropy of $H^1(\Xset)$ are not known in this setting.
@@
 
-For these reasons, analysis of least squares estimators for the nonparametric regression problem over Sobolev spaces typically assume $d = 1$---or more generally that $f_0 \in H^s(\Xset)$ for some $2s > d$.
-However, the Laplacian smoother $\wh{f}_{\LS}$ is a special least-squares estimator: it has a nice closed-form solution amenable to a bias-variance decomposition.
+% For these reasons, analysis of least squares estimators for the nonparametric regression problem over Sobolev spaces typically assume $d = 1$---or more generally that $f_0 \in H^s(\Xset)$ for some $2s > d$.
+However, the Laplacian smoother $\wh{f}_{\LS}$ is a special least-squares estimator---it has a nice closed-form solution amenable to a bias-variance decomposition.
 We therefore do not need to rely on empirical process theory, and so we will be able to analyze $\wh{f}_{\LS}$ and $T_{\LS}$ beyond the univariate setting.
 
+As we will see, the Sobolev embedding theorem has rather grave consequences for least-squares estimators over $H^1(\Xset)$ when $d > 1$, rendering such estimators ill-posed.
+However, the Laplacian smoother $\wh{f}_{\LS}$ is a special least-squares estimator.
+It has a nice closed-form solution amenable to a bias-variance decomposition, and so we will be able to analyze $\wh{f}_{\LS}$ and $T_{\LS}$ beyond the univariate setting.
+
 \paragraph{Connecting Laplacian smoothing to smoothing splines.} 
-To better motivate Laplacian smoothing as a tool for nonparametric regression, we review its connection to one of the most classical methods of nonparametric regression: smoothing splines.
-
-Smoothing splines are specially designed for regression over Sobolev spaces: they are penalized least squares estimators where the penalty is itself the Sobolev semi-norm.
-Mathematically, the smoothing spline estimator $\wh{f}_{\SM}$ is defined as the solution to the following optimization problem,
+To better motivate Laplacian smoothing as a tool for nonparametric regression, we connect it to one of the most classical methods of nonparametric regression: smoothing splines.
+
+Smoothing splines are specially designed for regression over Sobolev spaces, as they are penalized least squares estimators where the penalty is itself a Sobolev semi-norm.
+Mathematically, the (first-order) smoothing spline estimator $\wh{f}_{\SM}$ is defined as the solution to the following optimization problem,
 \begin{equation}
@@
 
-By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can tie the penalty used by the smoothing spline estimator $\wh{f}_{\SM}$ to that used by the Laplacian smoothing estimator $\wh{f}_{\LS}$.
+By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can tie the penalty used by the smoothing spline estimator $\wt{f}_{\SM}$ to that used by the Laplacian smoothing estimator $\wh{f}_{\LS}$.
 Suppose for the moment that $P$ is the uniform distribution over $\Xset$.
@@
 \end{equation}
-see \cite{ingster09} for a proof of this statement, and \cite{ariascastro2018} for a more extended discussion of the minimax paradigm in goodness-of-fit testing.
+see \cite{ingster09} for a proof of this statement, and \cite{ariascastro2018} for a more extended discussion of the minimax paradigm in nonparametric testing.
 When $d \geq 4$ we no longer have a continuous embedding of $H^1(\Xset)$ into $\Leb^4(\Xset)$; the functions in $H^1(\Xset)$ are very irregular, and the minimax rates in this regime are unknown.
 
-The estimator $\wh{f}_{\SM}$ and the level-$\alpha$ test $\phi_{\SM} := \1\{T_{\SM} \geq {t}_{\alpha}\}$ are known to be minimax rate-optimal---meaning they achieve the rates in~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius}, respectively---when $d = 1$ (see e.g.
+The estimator $\wt{f}_{\SM}$ and the level-$\alpha$ test $\wt{\phi}_{\SM} := \1\{T_{\SM} \geq \wt{t}_{\alpha}\}$ are known to be minimax rate-optimal---meaning they achieve the rates in~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius}, respectively---when $d = 1$ (see e.g.
 \cite{vandergeer2000} for the estimation result, and \cite{liu2019} for the testing result).
@@
 Finally, it is well known that if $\Xset$ satisfies the manifold hypothesis, meaning informally that it has intrinsic dimension $m < d$ (see~\ref{asmp:domain_manifold} in Section~\ref{sec:graph_sobolev_classes} for a formal statement), both~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$ \citep{bickel2007,ingster2000}.
-Encouragingly, the smoothing spline estimator $\wt{f}$ and test $\wt{\phi}$ achieve these faster rates, with one caveat: in order for smoothing splines to properly leverage the manifold hypothesis, the domain $\Xset$ must be known.
+Encouragingly, the smoothing spline estimator $\wt{f}_{\SM}$ and test $\wt{\phi}_{\SM}$ achieve these faster rates, with one caveat: in order for smoothing splines to properly leverage the manifold hypothesis, the domain $\Xset$ must be known.
 
 \paragraph{Consequences?} Intuitively, the connections between Laplacian smoothing and smoothing splines give us reason to hope that the strong statistical properties of the latter might be shared by the former.
-Unfortunately, it is not so easy to mathematically tie $\wh{f}$ to $\wt{f}$: for example, note that the convergence in~\eqref{eqn:seminorm_convergence} assumes $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:smoothing_spline} is either $H^1(\Xset)$ or $C^1(\Xset)$.
-Instead, we will take a different approach and analyze $\wh{f}$ by leveraging properties of the graph Sobolev class $H^1(G_{n,r})$, which we now introduce.
+Unfortunately, it is not so easy to mathematically tie $\wh{f}_{\LS}$ to $\wt{f}_{\SM}$: for example, note that the convergence in~\eqref{eqn:seminorm_convergence} assumes $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:smoothing_spline} is either $H^1(\Xset)$ or $C^1(\Xset)$.
+Instead, we will take a different approach and analyze $\wh{f}_{\LS}$ by leveraging properties of the graph Sobolev class $H^1(G_{n,r})$, which we now introduce.
 
@@
 Upper bounds on the error of smoothing splines rest on two crucial facts, both of which involve the continuum Laplace operator $\Delta_P$.
-The first is the \textit{a priori} assumption that the Sobolev semi-norm $|f|_{H^1(\Xset)}^2 = \langle f, \Delta_P f \rangle_{\Leb^2(P)} \leq M^2$.\footnote{The equality holding when $P$ is uniform and  only for functions $f \in C^2(\Xset)$ which satisfy boundary conditions.} The second is an upper bound on the metric entropy of $H^1(\Xset,M)$, characterized by the eigenvalue decay $\lambda_k(\Delta_P) \asymp k^{2/d}$ (i.e.
+The first is the \textit{a priori} assumption that the Sobolev semi-norm $|f_0|_{H^1(\Xset)}^2 = \langle f_0, \Delta_P f_0 \rangle_{\Leb^2(P)} \leq M^2$.\footnote{The equality holding when $P$ is uniform and  for functions $f_0 \in C^2(\Xset)$ which satisfy Dirichlet conditions.} The second is an upper bound on the metric entropy of $H^1(\Xset,M)$, characterized by the eigenvalue decay $\lambda_k(\Delta_P) \asymp k^{2/d}$ (i.e.
 Weyl's Law).
@@
 \paragraph{Graph Sobolev semi-norm.}
-We begin by upper bounding $f_0^T \Lap_{n,r} f_0$ under the assumption $f_0 \in H^1(\Xset,M)$. We will assume the kernel $K$ satisfies condition~\ref{asmp:kernel}.
+We begin by upper bounding $f^T \Lap_{n,r} f$ under the assumption $f_0 \in H^1(\Xset,M)$. We will assume the kernel $K$ satisfies condition~\ref{asmp:kernel}.
 \begin{enumerate}[label=(K\arabic*)]
@@
 	\label{prop:manifold_neighborhood_eigenvalue}
-	Assume~\ref{asmp:kernel}, ~\ref{asmp:domain_manifold}, and~\ref{asmp:radius2}. Then there exist constants $A_4, A_5$ and $a_4,a_5$ which do not depend on $n$ or $r$ such that 
+	Assume~\ref{asmp:kernel}, ~\ref{asmp:domain_manifold}, and~\ref{asmp:radius2}. It holds that
 	\begin{equation}
@@
 	\begin{equation*}
-	\frac{1}{n}\Bigl\|\wh{f}_{\LS} - f_0\Bigr\|_2^2 \leq \frac{C}{\delta} M^{d/(2 + d)} n^{-2/(2 + d)}
+	\Bigl\|\wh{f}_{\LS} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{d/(2 + d)} n^{-2/(2 + d)}
 	\end{equation*}
@@
 \end{theorem}
-To summarize: when $d = 1,2$ or $3$, the Laplacian smoothing estimator $\wh{f}_{\LS}$ has in-sample mean squared error with a constant factor of the minimax risk, except on a set of small probability.
+To summarize: when $d = 1,2$ or $3$, the Laplacian smoothing estimator $\wh{f}_{\LS}$ has in-sample mean squared error within a constant factor of the minimax risk, except on a set of small probability.
 Some remarks:
@@
 The upper bound in~\ref{asmp:radius1} is the largest value of $r$ such that this statement is still true.
-	\item The Laplacian smoothing estimator $\wh{f}_{\LS}$ is defined only at the design points $X_1,\ldots,X_n$ and we therefore measure error using the in-sample mean squared error $\frac{1}{n}\norm{\cdot}_2^2$.
-Because we work in the random design framework, the in-sample error is a random variable, and our bound is thus a bound in high probability.
+	\item The Laplacian smoothing estimator $\wh{f}_{\LS}$ is defined only at the design points $X_1,\ldots,X_n$ and we therefore measure error using the in-sample mean squared error $\norm{\cdot}_n^2$.
+Of course the in-sample error is a random variable, and our bound is thus a bound in high probability.
 That being said, it is possible to smoothly extend $\wh{f}_{\LS}$ to be defined on all of $\Xset$, or indeed all of $\Reals^d$---for instance, using the \emph{Nystrom extension}---and evaluate the error of the extension in $\Leb^2(\Xset)$ norm.
@@
 
-In Theorem~\ref{thm:laplacian_smoothing_testing}, we show this test also has small Type II error, uniformly over all $f_0$ separated from $0$ by at least the critical radius $\epsilon\bigl(H^1(\Xset,M)\bigr)$ given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
+Theorem~\ref{thm:laplacian_smoothing_testing} establishes that this test also has small Type II error, uniformly over all $f_0$ separated from $0$ by at least the critical radius $\epsilon\bigl(H^1(\Xset,M)\bigr)$ given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
 For this to hold, we will require a tight range of scalings for $r$.
@@
 Proving tight lower bounds over such classes is non-trivial, and to the best of our knowledge such an analysis remains outstanding.
-On the other hand, if we explicitly assume that $f \in \Leb^4(\Xset,M)$, then \cite{guerre02} show that the testing problem is characterized by the dimension-free lower bound $\epsilon^{\star}(\Leb^4(\Xset,M)) \gtrsim n^{-1/4}$.
+On the other hand, if we explicitly assume that $f \in \Leb^4(\Xset,M)$, then \cite{guerre02} show that the testing problem is characterized by the dimension-free lower bound $\epsilon^{2}(\Leb^4(\Xset,M)) \gtrsim n^{-1/2}$.
 Moreover, by training $\wh{f}_{\LS}$ to the interpolation limit, i.e.
@@
 	\end{equation}
-	for any $f_0 \in \Leb^4(\Xset,M)$ for which $\|f_0\|_{\Leb^2(\Xset)} \geq C bn^{-1/4}$. 
+	for any $f_0 \in \Leb^4(\Xset,M)$ for which $\|f_0\|_{\Leb^2(\Xset)}^2 \geq C b^2n^{-1/2}$. 
 	\item To compute the threshold $\wh{t}_b$, one must know each of the eigenvalues $\lambda_1(G_{n,r}),\ldots,\lambda_n(G_{n,r})$.

### 2020-10-07 notes/manuscripts/AISTATS/supplement_minimax_optimal_laplacian_smoothing.tex
Put AISTATS draft in AISTATS format; need to correct some overlapping lines.

--- 2020-10-06 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing2.tex
@@
-\documentclass{article}
+\documentclass[twoside]{article}
+\usepackage{aistats2021}
 \usepackage{amsmath}
@@
 \usepackage[shortlabels]{enumitem}
-\usepackage{fullpage}
 \usepackage{mathtools}
@@
 
-% \newcommand{\span}{\textrm{span}}
-
-\newtheoremstyle{alden}
-{6pt} % Space above
-{6pt} % Space below
-{} % Body font
-{} % Indent amount
-{\bfseries} % Theorem head font
-{.} % Punctuation after theorem head
-{.5em} % Space after theorem head
-{} % Theorem head spec (can be left empty, meaning `normal')
-
-\theoremstyle{alden}
-
-\newtheoremstyle{aldenthm}
-{6pt} % Space above
-{6pt} % Space below
-{\itshape} % Body font
-{} % Indent amount
-{\bfseries} % Theorem head font
-{.} % Punctuation after theorem head
-{.5em} % Space after theorem head
-{} % Theorem head spec (can be left empty, meaning `normal')
-
-\theoremstyle{aldenthm}
+%%% Theorem environments
 \newtheorem{theorem}{Theorem}
@@
 \begin{document}
-\title{Minimax optimal Laplacian smoothing}
-\author{Alden Green, Sivaraman Balakrishnan, and Ryan Tibshirani}
-\date{\today}
-\maketitle
-
-\begin{abstract}
-	\textcolor{red}{(TODO)}:
-\end{abstract}
-
-\section{Introduction}
-
-In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Reals^d$, and
-\begin{equation}
-\label{eqn:random_design_regression}
-Y_i = f_0(X_i) + \varepsilon_i
-\end{equation}
-with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise. We will assume the following throughout:
-\begin{enumerate}[label=(P\arabic*)]
-	\item
-	\label{asmp:bounded_lipschitz_density} 
-	$P$ admits a density $p$ bounded away from $0$ and $\infty$, i.e.
-	\begin{equation*}
-	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty,~~\textrm{for all $x \in \Xset$.}
-	\end{equation*}
-	Additionally, $p$ is Lipschitz, with Lipschitz constant $L_p$.
-\end{enumerate}
-Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present.
-
-In graph-based nonparametric regression, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$.
-The neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$.
-The $n \times n$ weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $\mathbf{W}_{ij}$ are given by
-\begin{equation*}
-\label{eqn:neighborhood_graph}
-{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}}{r}\Biggr).
-\end{equation*}
-where $\norm{\cdot} = \norm{\cdot}_{\Rd}$ is the Euclidean norm.
-Then the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$, and the graph Laplacian can be written as
-\begin{equation}
-\label{eqn:graph_Laplacian}
-\Lap_{n,r} = \bf{D} - \bf{W}
-\end{equation}
-
-The Laplacian smoothing estimator $\wh{f}_{\LS}$ \citep{smola2003} is a penalized least squares estimator, given by
-\begin{equation}
-\label{eqn:laplacian_smoothing}
-\wh{f}_{\LS} := \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_{n,r} f \biggr\}
-\end{equation}
-where $\rho > 0$ acts a tuning parameter on the penalty $f^T \Lap_{{n,r}} f$.
-Assuming~\eqref{eqn:laplacian_smoothing} is a reasonable estimator of $f_0$, the squared empirical norm
-\begin{equation}
-T_{\LS} = \frac{1}{n}\bigl\|\wt{f}_{\LS}\bigr\|_2^2 \label{eqn:laplacian_smoothing_test}
-\end{equation}
-is in turn a reasonable statistic to assess whether or not $f_0 = 0$.
-
-Laplacian smoothing has many advantages compared to other nonparametric regression methods. For instance:
-\begin{itemize}
-	\item It is fast, easy, and stable to compute.
-Letting ${\bf Y} = (Y_1,\ldots,Y_n) \in \Reals^n$, it is not hard to see that the minimizer $\wt{f}_{\LS} = (\rho \Lap_{n,r} + I)^{-1}{\bf Y}$ can be computed by solving a linear equation; in practice, one typically chooses $r$ to be small, with the result being that the matrix $\Lap_{n,r}$ is sparse.
-There exist known fast solvers of exactly this system.
-	\item Is is generalizable to non-standard data---for instance, text or images, or really any data modality on which it is possible to define a kernel.
-	\item It is easily adaptable to the semi-supervised learning setting. 
-\end{itemize}
-
-\paragraph{Related work.} For this reason a considerable body of work has emerged analyzing the properties of Laplacian smoothing, as well as of graph-based methods more generally.
-Roughly speaking this work can divided into two categories.
-
-Works in the first category view the responses $Y = \theta + \varepsilon$ as noisy samples of a regression function $\theta \in \Reals^n$ defined on the vertices of \emph{fixed graph} $G$.
-and bound the statistical error as a function of $\theta$ and $G$.
-(In the estimation literature, see for instance \citep{wang2016, sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}; in the testing literature, see \citep{sharpnack2013,sharpnack2013b,sharpnack2015}.) Note that in our setting, the graph $G_{n,r}$ and the evaluations $(f_0(X_1),\ldots,f_0(X_n))$ are random rather than fixed, and our assumptions are placed only on the underlying objects $P$ and $f_0$ from which these random object arise.
-As such, both our analysis and conclusions are naturally quite different.
-
-Works in the second category prove the consistency of neighborhood graph Laplacian smoothing and other neighborhood graph-based methods in the large-sample limit.
-They do so by establishing the convergence of the graph Laplacian $\Lap_{{n,r}}$ to the continuum Laplace operator $\Delta_P: C^2(\Xset) \to \Leb^2(\Xset)$,
-\begin{equation}
-\label{eqn:laplace_operator}
-\Delta_Pg := -\frac{1}{p} \dive(p^2 \nabla g)
-\end{equation}
-as $n \to \infty$ and $r \to 0$. This is meant in various senses:
-\begin{itemize}
-	\item Pointwise convergence, meaning that $\Lap_{n,r}f \to \Delta_Pf$ in an appropriate norm \citep{belkin03,belkin05,lafon04,hein05,singer06,gine06}. 
-	\item Spectral convergence, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$, where the latter are the Dirichlet eigenvalues of $\Delta_P$ \citep{belkin07,burago2014,trillos2018,trillos2019,calder2019}.\footnote{\label{footnote:dirichlet_eigenvalues} Formally, Dirichlet eigenvalues are solutions to the problem
-		$$
-		\Delta_Pg = \lambda g,\\
-		\restr{g}{\partial \Xset} = 0
-		$$} 
-	\item Convergence of norms, meaning $f^T \Lap_{n,r} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$ \citep{bousquet03,hein06,zhou11}. 
-\end{itemize}
-While these results imply \emph{consistency} of many different graph Laplacian based procedures, the statistical \emph{optimality} of Laplacian smoothing over neighborhood graphs is still not well understood.
-
-\paragraph{Our contributions.} Our three main contributions lie in this latter direction, and they can all be summarized as follows: Laplacian smoothing methods are minimax optimal over first order Sobolev spaces.
-In more detail, we will show that when $f_0$ belongs to the first order Sobolev ball $H^1(\Xset;M)$:
-\begin{itemize}
-	\item \textbf{Minimax optimal estimation.} With high probability, $\frac{1}{n}\bigl\|\wt{f}_{\LS} - f_0\bigr\|_{2}^2 \lesssim M^{d/(2+d)}n^{-2/(2 + d)}$.
-	\item \textbf{Minimax optimal testing.}
-	A level-$\alpha$ test constructed using $T_{\LS}$ has non-trivial power whenever $\norm{f_0}_{\Leb^2(\Xset)}^2 \gtrsim M^{2d/(4 + d)}n^{-4/(4 + d)}$. 
-	\item \textbf{Manifold adaptivity.}
-	If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$, both of the aforementioned rates hold with $d$ replaced by $m$.
-\end{itemize}
-
-\paragraph{Organization.} We now describe the structure of the rest of this paper, which builds to the aforementioned minimax results.
-\begin{itemize}
-	\item In Section~\ref{sec:minimax_optimal_regression_sobolev_spaces}, for completeness we review some standard results regarding minimax nonparametric regression over Sobolev spaces.
-	\item In Section~\ref{sec:graph_sobolev_classes}, we introduce the graph Sobolev class $H^1(G_{n,r})$, which we view as a graph-based analog to the continuum Sobolev class $H^1(\Xset)$.
-We show that key properties of $H^1(\Xset)$ used to derive minimax rates are shared by $H^1(G_{n,r})$.
-	\item In Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we leverage the results of Section~\ref{sec:graph_sobolev_classes} to derive tight upper bounds on the error of Laplacian smoothing methods.
-	\item In Section~\ref{sec:simulations} we provide some experiments supporting our theory, and conclude in Section~\ref{sec:discussion} with some potential methodological and theoretical extensions.
-\end{itemize}
-
-\textbf{Notation.}
-
-We use $A,A_1, \ldots$ and $C,C_1, \ldots$ to refer to constants which do not depend on $n$, $r$, or $f_0$; likewise with $a, a_1,\ldots$ and $c, c_1,\ldots$.
-
-\section{Minimax-optimal regression over Sobolev spaces}
-\label{sec:minimax_optimal_regression_sobolev_spaces}
-
-Before we get to our main results, we briefly review minimax optimal estimation and testing rates over Sobolev classes.
-We also single out a notable method which achieves these rates: \emph{smoothing splines}.
-As we will see, Laplacian smoothing can be seen as a graph-based---i.e.
-discrete---counterpart to the ``continuous-time'' approach of smoothing splines.
-
-\paragraph{Sobolev spaces, and the Sobolev embedding theorem.} We start by briefly introducing Sobolev spaces, emphasizing only the material relevant to our statistical context.
-
-Roughly speaking, (Hilbert-)Sobolev spaces contain functions $f \in \Leb^2(\Xset)$ with derivatives which themselves belong to $\Leb^2(\Xset)$.
-More formally, for a given integer $s$, the Sobolev space $H^{s}(\Xset)$ consists of all functions $f \in \Leb^2(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ exists and belongs to $\Leb^2(\Xset)$.
-For such functions, the Sobolev seminorm $\seminorm{f}_{H^{s}(\Xset)}$ and norm $\norm{f}_{H^{s}(\Xset)}$ are given by
-\begin{equation*}
-\seminorm{f}_{H^s(\Xset)}^2 = \sum_{\abs{\alpha} = s}\int_{\mathcal{X}} \Bigl|\bigl(D^{\alpha}f\bigr)(x)\Bigr|^2 \,dx, ~~ \norm{f}_{H^{s}(\Xset)}^2 = \sum_{k = 0}^{s} \seminorm{f}_{H^k(\Xset)}^2
-\end{equation*}
-and for a given $M > 0$, the order $s$ Sobolev ball is $H^s(\Xset, M) = \set{f: \norm{f}_{H^s(\Xset)} \leq M}$.
-The graph Sobolev seminorm $f^T \Lap_{n,r} f$ should be viewed as a random, discrete approximation to the continuum first-order Sobolev seminorm $|f|_{H^1(\Xset)}^2$, and so in our analysis of Laplacian smoothing we will mainly focus on the first-order Sobolev space $H^1(\Xset)$ hereafter.
-
-Much of statistical analysis of nonparametric regression over Sobolev spaces relies on the Sobolev embedding theorem, which establishes that functions in Sobolev spaces also satisfy other, (even more) classical notions of smoothness.
-Concretely, the Sobolev space $H^1(\Xset)$ will continuously embed either into a Holder space $C^{0,q}(\Xset)$ or an $\Leb^p$ space, depending on the dimension $d$.
-\begin{itemize}
-	\item When $d = 1$, any function $f \in H^1(\Xset)$ satisfies $\|f\|_{H^1(\Xset)} \leq C\|f\|_{C^{0,q}(\Xset)}$ for $q = 1/2.$
-	\item When $d > 2$, any function $f \in H^1(\Xset)$ satisfies $\|f\|_{H^1(\Xset)} \leq C\|f\|_{\Leb^{p}(\Xset)}$ for $p = 2d/(d - 2)$. 
-\end{itemize}
-In the above, the exponents $p$ and $q$ cannot be improved upon.
-
-When $d = 1$, the embedding of $H^1(\Xset)$ into $C^{0,q}(\Xset)$ facilitates a general analysis of (penalized) least estimators over $H^1(\Xset)$, using arguments from empirical process theory.
-However when $d \geq 2$, such arguments fall apart: for example, good bounds on the metric entropy of $H^1(\Xset)$ are not known in this setting.
-More troubling still, least squares estimators over $H^1(\Xset)$ are not even well-posed when $d \geq 2$.
-
-For these reasons, analysis of least squares estimators for the nonparametric regression problem over Sobolev spaces typically assume $d = 1$---or more generally that $f_0 \in H^s(\Xset)$ for some $2s > d$.
-However, the Laplacian smoother $\wh{f}_{\LS}$ is a special least-squares estimator: it has a nice closed-form solution amenable to a bias-variance decomposition.
-We therefore do not need to rely on empirical process theory, and so we will be able to analyze $\wh{f}_{\LS}$ and $T_{\LS}$ beyond the univariate setting.
-
-\paragraph{Connecting Laplacian smoothing to smoothing splines.} 
-To better motivate Laplacian smoothing as a tool for nonparametric regression, we review its connection to one of the most classical methods of nonparametric regression: smoothing splines.
-
-Smoothing splines are specially designed for regression over Sobolev spaces: they are penalized least squares estimators where the penalty is itself the Sobolev semi-norm.
-Mathematically, the smoothing spline estimator $\wh{f}_{\SM}$ is defined as the solution to the following optimization problem,
-\begin{equation}
-\label{eqn:smoothing_spline}
-\wt{f}_{\SM} = \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \cdot \seminorm{f}_{H^{1}(\Xset)}^2\biggr\}.
-\end{equation}
-In~\eqref{eqn:smoothing_spline}, the minimum can either be taken over all $f \in H^1(\Xset)$ or over all $f \in C^1(\Xset)$; we shall refer to the former as smoothing splines, and the latter as \emph{Lipschitz} smoothing splines.
-
-By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can tie the penalty used by the smoothing spline estimator $\wh{f}_{\SM}$ to that used by the Laplacian smoothing estimator $\wh{f}_{\LS}$.
-Suppose for the moment that $P$ is the uniform distribution over $\Xset$.
-Using the definition~\eqref{eqn:laplace_operator}, integrating by parts, and invoking the aforementioned convergence properties of $\Lap_{n,r}$, we have that for any $f \in C^{2}(\Xset)$ which satisfies Dirichlet boundary conditions (see footnote~\ref{footnote:dirichlet_eigenvalues}),
-\begin{equation}
-\label{eqn:seminorm_convergence}
-\abs{f}_{H^1(\Xset)}^2 = \int_{\Xset} \bigl(\Delta_P f\bigr)(x) \cdot f(x) \,dx = \lim_{n \to \infty} \frac{1}{n^{2}r^{(d + 2)}}f^T \Lap_{n,r} f.
-\end{equation}
-
-\eqref{eqn:seminorm_convergence} establishes asymptotic equality (up to a scaling factor) between the penalties in~\eqref{eqn:smoothing_spline} and~\eqref{eqn:laplacian_smoothing}, when the distribution $P$ is uniform.
-But it also sheds light on an additional feature of Laplacian smoothing, relevant when $P$ is non-uniform: the penalty $f^T \Lap_{n,r} f$ automatically adapts to the design density $p$, in a way that the penalty $|f|_{H^1(\Xset)}^2$ does not.
-This automatic adaptivity to $P$ turns out to be a formal advantage of Laplacian smoothing relative to smoothing splines, as we will see in Section~\ref{sec:minimax_optimal_laplacian_smoothing}.
-
-\paragraph{Minimax-optimality of smoothing splines.} 
-One reason all of this is intriguing is that, as mentioned previously, the smoothing spline estimator $\wt{f}_{\SM}$---and the resultant test statistic $T_{\SM} := \|\wt{f}_{\SM}\|_n^2$--- have strong statistical properties.
-To be precise, they achieve the well-known minimax rates over Sobolev spaces.
-
-Under~\ref{asmp:domain}, the minimax estimation rate over Sobolev balls is (see e.g. \citep{tsybakov2008_book})
-\begin{equation}
-\label{eqn:sobolev_space_estimation_minimax_rate}
-\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{2d/(2 + d)}n^{-2/(2 + d)}
-\end{equation}
-In the goodness-of-fit testing problem, we ask for a test function which can distinguish between the null and alternative hypotheses
-\begin{equation}
-\mathbf{H}_0: f_0 = 0, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{0\}
-\end{equation} 
-where $\mc{F}$ is some class of functions.
-The minimax critical radius is the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{F}_{\epsilon} := \mc{F} \cap \{f: \|f\|_{\Leb^2(\Xset)} \geq \epsilon\}$.
-Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing critical radius over $H^1(\Xset,M)$ is much smaller than the minimax estimation rate:
-\begin{equation}
-\label{eqn:sobolev_space_testing_critical_radius}
-\epsilon^2\Bigl(H^1(\Xset,M)\Bigr):= \inf\Biggl\{\epsilon^2: \inf_{\phi} \biggl[\Ebb_0[\phi] +  \sup_{f_0 \in H_{\epsilon}^1(\Xset,M)} \Ebb_{f_0}[1 - \phi]\biggr] \leq 2\alpha \Biggr\} \asymp M^{2d/(4 + d)}n^{-4/(4 + d)}~~\textrm{for $1 \leq d < 4$;}
-\end{equation}
-see \cite{ingster09} for a proof of this statement, and \cite{ariascastro2018} for a more extended discussion of the minimax paradigm in goodness-of-fit testing.
-When $d \geq 4$ we no longer have a continuous embedding of $H^1(\Xset)$ into $\Leb^4(\Xset)$; the functions in $H^1(\Xset)$ are very irregular, and the minimax rates in this regime are unknown.
-
-The estimator $\wh{f}_{\SM}$ and the level-$\alpha$ test $\phi_{\SM} := \1\{T_{\SM} \geq {t}_{\alpha}\}$ are known to be minimax rate-optimal---meaning they achieve the rates in~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius}, respectively---when $d = 1$ (see e.g.
-\cite{vandergeer2000} for the estimation result, and \cite{liu2019} for the testing result).
-When $d > 1$, as a consequence of the Sobolev Embedding Theorem\footnote{Or to be more precise, as a consequence of the \emph{tightness} of the Sobolev Embedding Theorem.} the variational problem~\eqref{eqn:smoothing_spline} is not well-posed: as illustrated in e.g.
-\cite{green93}, the criterion is minimized by interpolating the responses $\mathbf{Y}$, and the resulting estimator $\wt{f}_{\SM}$ will be inconsistent.
-By replacing the domain of minimization $H^1(\Xset)$ by $C^1(\Xset)$, we can force the solution to~\eqref{eqn:smoothing_spline} to be less wild, and \cite{birge1993} show that the resulting Lipschitz smoothing spline is not merely consistent but in fact minimax optimal when $d = 2$.\footnote{At least in the estimation problem.
-We could not find either a positive or negative result for the testing problem.} However when $d > 2$, the Holder space $C^1(\Xset)$ is itself too large, and even the Lipschitz smoothing spline has risk provably larger than optimal \cite{birge1993}.
-Previewing things a bit, we will see in Section~\ref{sec:minimax_optimal_laplacian_smoothing} that a similar phenomenon occurs for the Laplacian smoother $\wh{f}$, but interestingly the breakdown point will come when $d > 4$, rather than $d > 1$ or $d > 2$.
-
-Finally, it is well known that if $\Xset$ satisfies the manifold hypothesis, meaning informally that it has intrinsic dimension $m < d$ (see~\ref{asmp:domain_manifold} in Section~\ref{sec:graph_sobolev_classes} for a formal statement), both~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$ \citep{bickel2007,ingster2000}.
-Encouragingly, the smoothing spline estimator $\wt{f}$ and test $\wt{\phi}$ achieve these faster rates, with one caveat: in order for smoothing splines to properly leverage the manifold hypothesis, the domain $\Xset$ must be known.
-
-\paragraph{Consequences?} Intuitively, the connections between Laplacian smoothing and smoothing splines give us reason to hope that the strong statistical properties of the latter might be shared by the former.
-Unfortunately, it is not so easy to mathematically tie $\wh{f}$ to $\wt{f}$: for example, note that the convergence in~\eqref{eqn:seminorm_convergence} assumes $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:smoothing_spline} is either $H^1(\Xset)$ or $C^1(\Xset)$.
-Instead, we will take a different approach and analyze $\wh{f}$ by leveraging properties of the graph Sobolev class $H^1(G_{n,r})$, which we now introduce.
-
-\section{Graph Sobolev classes}
-\label{sec:graph_sobolev_classes}
-
-For $M > 0$, define the (first-order) graph Sobolev ball to be
-\begin{equation}
-\label{eqn:graph_sobolev_ball}
-H^1(G_{n,r},M) = \Bigl\{f \in \Reals^n: f^T \Lap_{n,r} f \leq M^2\Bigr\}
-\end{equation}
-where we recall that the \emph{graph Sobolev seminorm} $f^T \Lap_{{n,r}} f$ is precisely the penalty term in~\eqref{eqn:laplacian_smoothing}.
-
-To understand the role graph Sobolev balls play in our analysis of Laplacian smoothing, let us return for a moment to the continuous-time setting.
-Upper bounds on the error of smoothing splines rest on two crucial facts, both of which involve the continuum Laplace operator $\Delta_P$.
-The first is the \textit{a priori} assumption that the Sobolev semi-norm $|f|_{H^1(\Xset)}^2 = \langle f, \Delta_P f \rangle_{\Leb^2(P)} \leq M^2$.\footnote{The equality holding when $P$ is uniform and  only for functions $f \in C^2(\Xset)$ which satisfy boundary conditions.} The second is an upper bound on the metric entropy of $H^1(\Xset,M)$, characterized by the eigenvalue decay $\lambda_k(\Delta_P) \asymp k^{2/d}$ (i.e.
-Weyl's Law).
-In this section we will show that both these statements have analogues in the graph setting.
-\begin{itemize}
-	\item For functions $f$ belonging to $H^1(\Xset,M)$, the evaluations $f \in (f(X_1),\ldots,f(X_n)) \in \Reals^n$ are upper bounded,
-	\begin{equation}
-	\label{eqn:graph_sobolev_seminorm_vague}
-	f^T \Lap_{n,r} f \leq C M^2(n^{2}r^{(d + 2)}),
-	\end{equation}
-	\item For a range $1 \leq k \leq k_{\max}(n)$, the eigenvalues $\lambda_k(G_{n,r})$ satisfy
-	\begin{equation}
-	\label{eqn:neighborhood_graph_eigenvalue_vague}
-	ck^{2/d} \leq \frac{1}{nr^{d + 2}} \lambda_k(G_{n,r})  \leq Ck^{2/d}.
-	\end{equation}
-\end{itemize}
-Both statements will hold with high probability, and when we assume $\Xset$ is a manifold of intrinsic dimension $m$, they will hold with $d$ replaced by $m$.
-These results elucidate why Laplacian smoothing ``works'' as well as its continuous-time counterpart ---in the sense of being statistically rate-optimal over (continuum) Sobolev classes.
-
-\paragraph{Graph Sobolev semi-norm.}
-We begin by upper bounding $f_0^T \Lap_{n,r} f_0$ under the assumption $f_0 \in H^1(\Xset,M)$. We will assume the kernel $K$ satisfies condition~\ref{asmp:kernel}.
-\begin{enumerate}[label=(K\arabic*)]
-	\item
-	\label{asmp:kernel}
-	$K:[0,\infty) \to [0,\infty)$ is a non-increasing function supported on $[0,1]$, its restriction to $[0,1]$ is Lipschitz, and $K(1) > 0$. Additionally, it is normalized so that
-	\begin{equation*}
-	\int_{\Reals^d} K\bigl(\norm{z}\bigr) \,dz = 1.
-	\end{equation*}
-	Let $\sigma_K := \frac{1}{d} \int_{\Rd} \|x\|^2 K(\|x\|) \,dx$.
-\end{enumerate}
-This is a relatively mild assumption: the choice of kernel is under the user's control, and moreover~\ref{asmp:kernel} covers many (though certainly not all) common kernel choices.
-In Lemma~\ref{lem:graph_sobolev_seminorm}, we prove that if the density $p$ is merely upper bounded then~\eqref{eqn:graph_sobolev_seminorm_vague} holds with respect to the graph Sobolev semi-norm.
-\begin{lemma}
-	\label{lem:graph_sobolev_seminorm}
-	Assume~\ref{asmp:kernel} and additionally that $p(x) \leq p_{\max}$ for all $x \in \Xset$. Then, for any $f \in H^1(\Xset)$, 
-	\begin{equation}
-	\label{eqn:graph_sobolev_seminorm}
-	f^T \Lap_{n,r} f \leq \frac{p_{\max}^2 \sigma_K}{\delta} n^2 r^{d + 2} |f|_{H^1(X)}^2
-	\end{equation}
-	with probability at least $1 - \delta$.
-\end{lemma}
-A few remarks:
-\begin{itemize}
-	\item 
-	The inequality~\eqref{eqn:graph_sobolev_seminorm} resembles~\eqref{eqn:seminorm_convergence}.
-However, as we pointed out previously, to derive~\eqref{eqn:seminorm_convergence} one needs $f \in C^{2}(\Xset)$ as opposed to $f \in H^1(\Xset)$, and imposing this extra degree of regularity is unsuitable for our purposes.
-We use a more careful analysis to prove Lemma~\ref{lem:graph_sobolev_seminorm}, but note well that~\eqref{eqn:graph_sobolev_seminorm} is only an upper bound rather than a statement about consistency.
-This is the price we pay for insisting that $f$ have only one derivative; fortunately an upper bound will be sufficient for our purposes in Section~\ref{sec:minimax_optimal_laplacian_smoothing}.
-	\item We note also that~\eqref{eqn:graph_sobolev_seminorm}---which holds with probability $1 - \delta$---depends proportionally on $1/\delta$, as opposed to $\log(1/\delta)$.
-The reason: the Sobolev assumption $f \in H^1(\Xset)$ only grants us control over the second moment of $f$ and its first derivative, and so we are afforded only the relatively weak concentration implied by Markov's inequality.
-If we assume $f$ and its first derivative are more uniformly bounded--for instance if $f \in C^1(\Xset)$---we can use Hoeffding's inequality to get sharper upper bounds.
-\end{itemize}
-
-\paragraph{Graph Laplacian eigenvalues.} Lemma~\ref{lem:neighborhood_eigenvalue} establishes upper and lower bounds on the graph Laplacian eigenvalues $\lambda_{k}(G_{n,r})$ in terms of the continuum Laplace eigenvalues $\lambda_k(\Delta_P)$.
-These bounds hold with high probability under appropriate regularity conditions on the domain $\Xset$, and for a range of radii $r$.
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{1}
-	\item
-	\label{asmp:domain}
-	The domain $\Xset$ is an open, connected set with Lipschitz boundary.
-\end{enumerate}
-\begin{enumerate}[label=(R\arabic*)]
-	\item
-	\label{asmp:radius1}
-	The radius $r$ satisfies $(\log n/n)^{1/d} \leq r \leq c_1$.
-\end{enumerate}
-\begin{lemma}
-	\label{lem:neighborhood_eigenvalue}
-	Suppose the density $p$, the domain $\Xset$, the kernel $K$ and the radius $r$ satisfy~\ref{asmp:bounded_lipschitz_density}, \ref{asmp:domain}, \ref{asmp:kernel},and~\ref{asmp:radius1} respectively.
-The following statement holds with probability at least $1 - A_1n\exp(-a_1nr^{d})$: for any $\ell \in \mathbb{N}$ such that
-	\begin{equation}
-	\label{eqn:neighborhood_eigenvalue_1}
-	1 - A_1 r \sqrt{\lambda_{\ell}(\Delta_P)} \geq \frac{1}{2}
-	\end{equation}
-	we have that
-	\begin{equation}
-	\label{eqn:eigenvalue_bound}
-	a_1 \lambda_k(G_{n,r}) \leq nr^{d+2} \lambda_k(\Delta_P) \leq A_1 \lambda_k(G_{n,r}),~~\textrm{for all $1 \leq k \leq \ell$}
-	\end{equation}
-\end{lemma}
-
-The scaling~\eqref{eqn:neighborhood_graph_eigenvalue_vague} then follows by Weyl's Law (see Lemma 28 of~\citep{dunlop2020} for a proof that~\ref{asmp:bounded_lipschitz_density} and~\ref{asmp:domain} are sufficient for Weyl's Law to hold.).
-\begin{corollary}
-	\label{cor:neighborhood_eigenvalue}
-	Suppose the density $p$ and the domain $\Xset$ satisfy~\ref{asmp:bounded_lipschitz_density} and~\ref{asmp:domain}; then there exist constants $a_2$ and $A_2$ such that
-	\begin{equation}
-	\label{eqn:weyls_law}
-	a_2k^{2/d} \leq \lambda_k(\Delta_P) \leq A_2k^{2/d}~~\textrm{for all $k \in \mathbb{N}, k > 1$}.
-	\end{equation}
-	Therefore under the conditions of Lemma~\ref{lem:neighborhood_eigenvalue}, we have that
-	\begin{equation}
-	\label{eqn:neighborhood_eigenvalue_2}
-	a_3 \cdot \min\Bigl\{nr^{d + 2}k^{2/d},nr^d\Bigr\} \leq \lambda_k(G_{n,r}) \leq A_3 \cdot \min\Bigl\{nr^{d + 2}k^{2/d}, nr^d\Bigr\}~~\textrm{for all $2 \leq k \leq n$}
-	\end{equation}
-	with probability at least $1 - A_1n\exp(-a_1nr^d)$.
-\end{corollary}
-
-We see that the inequalities in~\eqref{eqn:neighborhood_eigenvalue_2} match the desired scaling~\eqref{eqn:neighborhood_graph_eigenvalue_vague} for all $2 \leq k \leq (1/r)^d$.
-Indeed, if we could choose $r = n^{-1/d}$, we would get the desired scaling for all $2 \leq k \leq n$, but unfortunately choosing $r$ to be so small would violate~\ref{asmp:radius1}.
-Nevertheless, by choosing $r$ to be relatively small while still being in accordance with~\ref{asmp:radius1}, we can guarantee that only some of the largest eigenvalues $\lambda_k(G_{n,r})$ will fail to satisfy~\eqref{eqn:neighborhood_graph_eigenvalue_vague}.
-Fortunately, the behavior of these large eigenvalues will not meaningfully contribute to the error of $\wh{f}_{\LS}$ or $T_{\LS}$, as we shall see.
-
-\paragraph{Manifold hypothesis.}
-
-As mentioned in the introduction, an attractive quality of the neighborhood graph $G_{n,r}$ is that it captures the geometry of $P$ and $\Xset$.
-We now review two manifestations of this phenomenon, from \cite{trillos2019} and \cite{calder2019}: both the graph Sobolev semi-norm and graph Laplacian eigenvalues automatically adapt to the intrinsic dimension of $\Xset$.
-Formally, we make the following assumption on $\Xset$ (in order to directly use the results of~\cite{trillos2019} and~\cite{calder2019}).
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{2}
-	\item 
-	\label{asmp:domain_manifold}
-	The measure $P$ is supported on a compact, connected, smooth submanifold $\Xset$ of $\Reals^d$; the domain $\Xset$ is of fixed dimension $1 \leq m \leq d$ and without boundary.
-Moreover, $P$ has a density $p$ with respect to the volume form of $\Xset$ which is bounded away from $0$ and $\infty$,
-	\begin{equation*}
-	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty
-	\end{equation*}
-	for all $x \in \Xset$, and the density $p$ is Lipschitz with Lipschitz constant $L_p$.
-\end{enumerate}
-\begin{enumerate}[label=(R\arabic*)]
-	\setcounter{enumi}{1}
-	\item 
-	\label{asmp:radius2}
-	The radius $r$ satisfies $(\log n/n)^{1/m} \leq r \leq c_2$.
-\end{enumerate}
-\begin{proposition}[Lemma 5 of~\cite{trillos2019}]
-	\label{prop:manifold_graph_sobolev_seminorm}
-	Assume~\ref{asmp:kernel},~\ref{asmp:domain_manifold} and~\ref{asmp:radius2}. Then for any $f \in H^1(\Xset)$,
-	\begin{equation*}
-	f^T \Lap_{n,r} f \leq \frac{(1 + L_pr)^2}{\delta}n^{2}r^{m + 2} \sigma_K |f|_{H^1(\Xset)}^2
-	\end{equation*}
-	with probability at least $1 - \delta$.
-\end{proposition}
-
-\begin{proposition}[Restatement of Theorem 4 of~\cite{trillos2019}]
-	\label{prop:manifold_neighborhood_eigenvalue}
-	Assume~\ref{asmp:kernel}, ~\ref{asmp:domain_manifold}, and~\ref{asmp:radius2}. Then there exist constants $A_4, A_5$ and $a_4,a_5$ which do not depend on $n$ or $r$ such that 
-	\begin{equation}
-	\label{eqn:manifold_neighborhood_eigenvalue}
-	a_5 \cdot \min\Bigl\{nr^{m + 2}k^{2/m},nr^d\Bigr\} \leq \lambda_k(G_{n,r}) \leq A_5 \cdot \min\Bigl\{nr^{m + 2}k^{2/m}, nr^m\Bigr\}~~\textrm{for all $2 \leq k \leq n$}	
-	\end{equation}
-	with probability at least $1 - A_4n\exp(-a_4nr^m)$.
-\end{proposition}
-Note that~\ref{asmp:domain_manifold} assumes $\Xset$ is without boundary, and so Propositions~\ref{prop:manifold_graph_sobolev_seminorm} and~\ref{prop:manifold_neighborhood_eigenvalue} do not imply Lemmas~\ref{lem:graph_sobolev_seminorm} or~\ref{lem:neighborhood_eigenvalue}.
-
-\section{Minimax optimality of Laplacian smoothing}
-\label{sec:minimax_optimal_laplacian_smoothing}
-
-We now leverage the results of Section~\ref{sec:graph_sobolev_classes} to formalize the main conclusions of this work: that Laplacian smoothing methods are minimax rate-optimal.
-
-For all results in this section, we will assume that the density~$p$ satisfies~\ref{asmp:bounded_lipschitz_density} and the kernel $K$ satisfies~\ref{asmp:kernel}.
-Additionally, within our theorem statements we will use $c$ and $C$ to denote constants which possibly differ from theorem to theorem, and which may depend on $K$, $d$, $p$, and $\Xset$ but do not depend on $f_0$, the radius of the Sobolev ball $M$, or the sample size $n$.
-
-\paragraph{Estimation error of Laplacian smoothing.} 
-Suppose the domain $\Xset$ satisfies~\ref{asmp:domain}.
-Under appropriate conditions, the Laplacian smoothing estimator $\wh{f}_{\LS}$ achieves, up to constants, the minimax risk $n^{-2/(2 + d)}$ over all $f_0 \in H^1(\Xset,M)$.
-To prove this, we will invoke Lemma~\ref{lem:graph_sobolev_seminorm} and Corollary~\ref{cor:neighborhood_eigenvalue} to suitably upper bound the squared-bias and variance of $\wt{f}_{\LS}$, respectively.
-In order to do so, we must restrict the range of the neighborhood graph radius $r$.
-\begin{enumerate}[label=(R\arabic*)]
-	\setcounter{enumi}{2}
-	\item 
-	\label{asmp:ls_kernel_radius_estimation}
-	The neighborhood graph radius $r$ satisfies
-	\begin{equation*}
-	\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq r \leq n^{-3/(4 + 2d)} M^{(d - 4)/(4 + 2d)}.
-	\end{equation*}
-\end{enumerate}
-\begin{theorem}
-	\label{thm:laplacian_smoothing_estimation1}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d < 4$ and $f_0 \in H^1(\Xset,M)$.
-Suppose that that the neighborhood graph $G_{n,r}$ is computed with a radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_estimation},  and the Laplacian smoothing estimator $\wh{f}_{\LS}$ with $\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$.
-It holds that
-	\begin{equation*}
-	\frac{1}{n}\Bigl\|\wh{f}_{\LS} - f_0\Bigr\|_2^2 \leq \frac{C}{\delta} M^{d/(2 + d)} n^{-2/(2 + d)}
-	\end{equation*}
-	with probability at least $1 - \delta -  A_1n\exp(-a_1nr^d) - \exp(-c M^{d/(2d + 4)} n^{d/(2+d)})$.
-\end{theorem}
-To summarize: when $d = 1,2$ or $3$, the Laplacian smoothing estimator $\wh{f}_{\LS}$ has in-sample mean squared error with a constant factor of the minimax risk, except on a set of small probability.
-Some remarks:
-\begin{itemize}
-	\item When $d = 4$, our analysis results in an upper bound on the mean squared error within a $\log n$ factor of the minimax risk, but when $d \geq 5$ our analysis yields upper bounds that are much worse than the minimax rate.
-This mirrors the conclusions of~\cite{sadhanala16}, who investigate estimation rates of Laplacian smoothing over the $d$-dimensional grid graph.
-In our setting, when $d = 4$ we would set $r \asymp (\log n/n)^{1/4}$---slightly outside the scaling~\ref{asmp:ls_kernel_radius_estimation}---to get the near-optimal rate.
-	\item The range of radii $r$ imposed by~\ref{asmp:ls_kernel_radius_estimation} is  compatible with practice, where by far the most common choice of radius is the connectivity threshold $r \asymp (\log(n)/n)^{1/d}$, chosen to make the graph $G_{n,r}$ as sparse as possible while still being connected.
-	\item The tuning parameter $\rho$ is chosen to balance the (conditional on ${\bf X}$) squared-bias and variance, i.e. chosen as large as possible while still with high probability satisfying
-	\begin{equation}
-	\label{eqn:laplacian_smoothing_bias_variance}
-	\frac{\rho}{n} f_0^T \Lap_{{n,r}} f_0 \lesssim \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k}(G_{n,r}) + 1\bigr)^2}.
-	\end{equation}
-	To balance bias and variance as in~\eqref{eqn:laplacian_smoothing_bias_variance} we require control on all the graph eigenvalues $\lambda_1(G_{n,r}),\ldots,\lambda_n(G_{n,r})$.
-As mentioned in our commentary after Lemma~\ref{lem:neighborhood_eigenvalue}, when $r$ is small enough the large eigenvalues $\lambda_k(G_{n,r})$ contribute negligibly to the variance term.
-The upper bound in~\ref{asmp:radius1} is the largest value of $r$ such that this statement is still true.
-	\item The Laplacian smoothing estimator $\wh{f}_{\LS}$ is defined only at the design points $X_1,\ldots,X_n$ and we therefore measure error using the in-sample mean squared error $\frac{1}{n}\norm{\cdot}_2^2$.
-Because we work in the random design framework, the in-sample error is a random variable, and our bound is thus a bound in high probability.
-That being said, it is possible to smoothly extend $\wh{f}_{\LS}$ to be defined on all of $\Xset$, or indeed all of $\Reals^d$---for instance, using the \emph{Nystrom extension}---and evaluate the error of the extension in $\Leb^2(\Xset)$ norm.
-Assuming that the extension of our estimators and the regression function $f_0$ are suitably smooth, tools from empirical process theory will guarantee that the $\Leb^2(\Xset)$ error is not too much greater than the in-sample error, but we do not further pursue the details here.
-\end{itemize}
-
-\paragraph{Testing error of Laplacian smoothing.}
-Let us define a test using the statistic $T_{\LS}$. For $b \geq 1$, define the threshold $\wh{t}_b$ to be
-\begin{equation*}
-\wh{t}_{b} := \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k(G_{n,r}) + 1\bigr)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k(G_{n,r}) + 1\bigr)^4}},
-\end{equation*}
-and the Laplacian smoothing test $\wh{\phi}_{\LS} := \1\bigl\{T_{\LS} \leq \wh{t}_b\bigr\}$.
-(Clearly the test $\wh{\phi}_{\LS}$ depends on $b$, but we suppress this notationally.) The number $b \geq 1$ is selected by the user, with the choice made based on the tolerated level of Type I and Type II error.
-In fact, the threshold $\wh{t}_b$ is precisely the right choice to control the Type I error of $\wh{\phi}_{\LS}$; as we show in Lemma~\ref{lem:ls_fixed_graph_testing} in the supplementary material,
-\begin{equation*}
-\Ebb_0\Bigl[\wh{\phi}_{\LS}\Bigr] \leq \frac{1}{b^2}.
-\end{equation*}
-
-In Theorem~\ref{thm:laplacian_smoothing_testing}, we show this test also has small Type II error, uniformly over all $f_0$ separated from $0$ by at least the critical radius $\epsilon\bigl(H^1(\Xset,M)\bigr)$ given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
-For this to hold, we will require a tight range of scalings for $r$.
-\begin{enumerate}[label=(R\arabic*)]
-	\setcounter{enumi}{3}
-	\item 
-	\label{asmp:ls_kernel_radius_testing}
-	The neighborhood graph radius $r$ satisfies
-	\begin{equation*}
-	\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq r \leq M^{(16 - d)/(8 + 2d)}n^{(d - 20)/(32 + 8d)}.
-	\end{equation*}
-\end{enumerate}
-
-\begin{theorem}
-	\label{thm:laplacian_smoothing_testing}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d < 4$ and $f_0 \in H^1(\Xset,M)$ for $M \leq n^{(4 - d)/(4 + d)}$.
-Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}, and the Laplacian smoothing test $\wh{\phi}_{\LS}$ with $\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$, and $b \geq 1$.
-For all $f_0$ such that
-	\begin{equation}
-	\label{eqn:laplacian_smoothing_testing}
-	\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2d/(4 + d)} n^{-4/(4 + d)}
-	\end{equation} 
-	the Type II error is upper bounded:
-	\begin{equation*}
-	\Ebb_{f_0}\Bigl[1 - \wh{\phi}_{\LS} \Bigr] \leq C\biggl(\frac{1}{b}\Bigl[1 + n^{-d(4 + d)}M^{-2d/(4 + d)}\Bigr] + A_1n\exp\bigl(-a_1nr^d\bigr)\biggr).
-	\end{equation*}
-\end{theorem}
-
-A few remarks:
-\begin{itemize}
-	\item As mentioned in Section~\ref{sec:minimax_optimal_regression_sobolev_spaces}, when $d \geq 4$ the Sobolev balls $H^1(\Xset,M)$ include quite irregular functions $f \not\in \Leb^4(\Xset)$.
-Proving tight lower bounds over such classes is non-trivial, and to the best of our knowledge such an analysis remains outstanding.
-On the other hand, if we explicitly assume that $f \in \Leb^4(\Xset,M)$, then \cite{guerre02} show that the testing problem is characterized by the dimension-free lower bound $\epsilon^{\star}(\Leb^4(\Xset,M)) \gtrsim n^{-1/4}$.
-Moreover, by training $\wh{f}_{\LS}$ to the interpolation limit, i.e.
-setting $\rho = 0$, the test $\wh{\phi}_{\LS}$ achieves (up to constants) this lower bound.
-That is,
-	\begin{equation}
-	\label{eqn:laplacian_smoothing_testing_low_smoothness}
-	E_{f_0}\Bigl[1 - \wh{\phi}_{\LS}\Bigr] \leq \frac{C(1 + M^4)}{b^2}
-	\end{equation}
-	for any $f_0 \in \Leb^4(\Xset,M)$ for which $\|f_0\|_{\Leb^2(\Xset)} \geq C bn^{-1/4}$. 
-	\item To compute the threshold $\wh{t}_b$, one must know each of the eigenvalues $\lambda_1(G_{n,r}),\ldots,\lambda_n(G_{n,r})$.
-Computing all $n$ of these eigenvalues is far more expensive than computing $T_{\LS}$.
-That being said, in practice we would not recommend using $\wh{t}_b$ anyway: the threshold is justified by asymptotic theory, but in practice $n$ is always finite, and using such thresholds can result in substantial loss of efficiency.
-We would instead make the standard recommendation to calibrate via permutation.
-\end{itemize}
-It is worth circling back to our comparison of Laplacian smoothing methods to smoothing splines.
-While we motivated Laplacian smoothing methods in part by suggesting they might replicate the statistical properties of smoothing splines, we have in fact managed to show something stronger.
-Theorems~\ref{thm:laplacian_smoothing_estimation1} and~\ref{thm:laplacian_smoothing_testing} imply that Laplacian smoothing methods are minimax optimal for both estimation and testing in dimensions $d = 1,2$ and $3$, and are nearly minimax-optimal for the estimation problem when $d = 4$.
-On the other hand, as covered in Section~\ref{sec:minimax_optimal_regression_sobolev_spaces} smoothing spline methods are optimal only when $d = 1$ or, for Lipschitz smoothing splines, $d = 2$.
-Thus, we see a somewhat surprising gap emerge between the optimality properties of the discrete-time (Laplacian smoothing) and continuous-time (smoothing spline) approaches, in favor of the former.
-
-\paragraph{Manifold adaptivity.}
-
-In Section~\ref{sec:graph_sobolev_classes}, we reviewed that when $\Xset$ is a submanifold of $\Rd$, the graph Laplacian automatically adapts to the geometry---and in particular, the intrinsic dimension $m$---of $\Xset$.
-This translates into improved upper bounds on the estimation and testing error of Laplacian smoothing; the upper bounds now depend on $m$ instead of $d$.
-
-\begin{theorem}
-	\label{thm:laplacian_smoothing_manifold}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
-Suppose that $\Xset$ satisfies~\ref{asmp:domain_manifold} for $m < 4$, and that $f_0 \in H^1(\Xset,M)$.
-For appropriate choices of tuning parameters $r$, $\rho$, and $b \geq 1$, each of the following statements holds.
-	\begin{itemize}
-		\item The Laplacian smoothing estimator satisfies
-		\begin{equation*}
-		\frac{1}{n}\bigl\|\wh{f}_{\LS} - f_0\bigr\|_2^2 \leq \frac{C}{\delta}M^{m/(2 + m)}n^{-2/(2 + m)}
-		\end{equation*}
-		with probability at least $1 - \delta - A_4n\exp(-a_4nr^m) - \exp(-c M^{m/(2m + 4)} n^{m/(2+m)})$.
-		\item Suppose $M \leq n^{(4 - m)/(4 + m)}$. For all $f_0$ such that
-		\begin{equation*}
-		\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2m/(4 + m)} n^{-4/(4 + m)}
-		\end{equation*} 
-		the type II error of $\wh{\phi}$ is upper bounded:
-		\begin{equation*}
-		\Ebb_{f_0}\Bigl[1 - \wh{\phi}_{\LS} \Bigr] \leq C\biggl(\frac{1}{b}\Bigl[1 + n^{-m(4 + m)}M^{-2m/(4 + m)}\Bigr] + A_4n\exp\bigl(-a_4nr^m\bigr)\biggr).
-		\end{equation*}
-	\end{itemize}
-\end{theorem}
-In our supplementary material, we specify the right choices of $\rho$, as well as the permissible ranges of $r$.
-For now, we merely point out that they will depend only on the intrinsic dimension $m$, and not on any other properties of the manifold $\Xset$.
-In all other respects, $\wh{f}_{\LS}$ and $T_{\LS}$ are computed without any knowledge of $\Xset$.
-By contrast, the penalty in~\eqref{eqn:smoothing_spline} must be specially tailored to $\Xset$ in order for $\wt{f}_{\SM}$ and $T_{\SM}$ to achieve the minimax estimation and testing rates.
-This reveals another advantage of Laplacian smoothing as opposed to smoothing splines.
-
-\section{Simulations}
-\label{sec:simulations}
-
-\textcolor{red}{(TODO)}
-
-\section{Discussion}
-\label{sec:discussion}
-
-\textcolor{red}{(TODO)}
-
-A few discussion items:
-\begin{itemize}
-	\item \textbf{Higher-order smoothness classes.} Laplacian smoothing can also be adapted to take advantage of additional assumed regularity on $f_0$, i.e.
-$f_0 \in H^s(\mc{\Xset})$ for $s > 1$.
-In the very special case of $s = 2$ and $d = 2$, we can show that this adapted estimator achieves the sharper minimax rates over these classes, but the general story--for all combinations of $s$ and $d$---remains beyond our reach.
-	\item \textbf{Possible extensions}. Methodologically: kNN graphs and different normalizations of the Laplacian. Theoretically: relaxing~\ref{asmp:bounded_lipschitz_density}. 
-\end{itemize}
-
-\clearpage
-
-\appendix
-
-\clearpage
+
+\onecolumn
+\aistatstitle{Supplementary materials for: \\
+	Minimax Optimal Laplacian Smoothing}
+
+\section{Organization of Supplement}
 
@@
 
-\clearpage
-
-\bibliographystyle{plainnat}
-\bibliography{../../../graph_testing_bibliography}
-
 \end{document}

### 2020-10-07 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing3.tex
Begun a 2nd attempt at AISTATS paper on Laplacian smoothing.

\documentclass[twoside]{article}
\usepackage{aistats2021}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{mathtools}
\usepackage{tikz}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\iid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}
\newcommand{\lambdavec}{\boldsymbol{\lambda}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}
\newcommand{\Lap}{L}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}
\newcommand{\Id}{I}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Vset}{\mathcal{V}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\SM}{\mathrm{SM}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

%%% Theorem environments
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}

\twocolumn[

	\aistatstitle{Minimax optimal Laplacian smoothing}

	\aistatsauthor{Alden Green, Sivaraman Balakrishnan, and Ryan J. Tibshirani}

	\aistatsaddress{Carnegie Mellon University}]

\begin{abstract}
	In this paper we study the statistical properties of Laplacian smoothing, a graph-based approach to non-parametric regression. We show that under standard regularity conditions, the Laplacian smoothing estimator $\wh{f}$, and a test using the statistic $T = \frac{1}{n}\|\wh{f}\|_2^2$, attain the minimax optimal estimation and testing rates of convergence over the first-order Sobolev class $H^1(\Xset)$, where $\Xset \subset \Rd$ and $1 \leq d \leq 4$. Additionally, we prove that Laplacian smoothing is manifold adaptive: if $\Xset$ is an $m$ dimensional manifold embedded in $\Reals^d$ with $m \ll d$, then the error rate of Laplacian smoothing methods depends only on the intrinsic dimension $m$ and not on the ambient dimension $d$. We support our theory with simulations.
\end{abstract}

\section{INTRODUCTION}

Suppose we observe responses according to the signal plus noise model:
\begin{equation}
\label{eqn:signal_plus_noise_model}
Y_i = f_{0,i} + \varepsilon_i
\end{equation}
with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise. Our goal is to perform statistical inference on the unknown regression function $f_0$, by which we mean either \emph{estimating} $f_0$ or more simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present.

The Laplacian smoothing estimator $\wh{f}$ \citep{smola2003} is a penalized least squares estimator, defined on a graph. Letting $G = \bigl(\{1,\ldots,n\},\mathbf{W}\bigr)$ be a weighted graph over the vertices $\{1,\ldots,n\}$, 
\begin{equation}
\label{eqn:laplacian_smoothing}
\wh{f} :=  \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap f \biggr\},
\end{equation}
where $\Lap$ is the graph Laplacian matrix (defined in Section 2), and the penalty
\begin{equation*}
f^T \Lap f = \frac{1}{2} \sum_{i,j = 1}^{n} {\bf W}_{ij}\bigl(Y_i - Y_j\bigr)^2
\end{equation*}
encourages $\wh{f}_i \approx \wh{f}_j$ when ${\bf W}_{ij}$ is large. Assuming~\eqref{eqn:laplacian_smoothing} is a reasonable estimator of $f_0$, the squared empirical norm
\begin{equation}
\label{eqn:laplacian_smoothing_test}
\wh{T} = \frac{1}{n}\bigl\|\wh{f}\bigr\|_2^2 
\end{equation}
is in turn a reasonable statistic to assess whether or not $f_0 = 0$.

Of course there exist many methods for nonparametric regression, but Laplacian smoothing has its own set of advantages. For instance:
\begin{itemize}
	\item \emph{Computational ease.} Laplacian smoothing is fast, easy, and stable to compute. The estimate $\wh{f}$ can be computed by solving a symmetric, diagonally dominant linear system. There are known extremely fast and reliable solvers for exactly this problem \textcolor{red}{(TODO)}: references.
	\item \emph{Generalizability.} Laplacian smoothing is well defined whenever one can associate a graph with some observed responses. This generality lends itself to many different data modalities: for instance sentiment analysis and image classification, to name only two. \textcolor{red}{(TODO)}: references
	\item \emph{Supervision.} Although we study Laplacian smoothing in the supervised learning setup~\eqref{eqn:signal_plus_noise_model}, it can be easily adapted to the semi-supervised or unsupervised settings.
\end{itemize}

For these reasons, a body of work has emerged analyzing the statistical properties of Laplacian smoothing, and graph-based methods more generally. Roughly speaking, these works can be divided into two categories, based on the perspective they adopt. 
\begin{itemize}
	\item In the \emph{discrete (fixed design) perspective}, both $G$ and $f_0 \in \Reals^n$ are treated as fixed, and the regression ``function'' is really a vector. In this context, tight upper bounds have been derived on the error of graph-based estimators, which certify that such estimators are \emph{optimal} over ``function'' classes $\mc{F} \subseteq \Reals^n$. The downside of this work is that, depending on the context, it may be unnatural to assume a priori that the graph $G$ is fixed, and that the signal $f_0$ belongs to some discrete ``function'' class.
	\item In the arguably more natural \emph{continuum (random design) perspective}, in addition to the responses $Y_i$ one observes randomly sampled design points $\{X_1,\ldots,X_n\}$ belonging to a domain $\Xset \subseteq \Rd$, and assumes that $f_{0,i} \equiv f_0(X_i)$ are $n$ evaluations of a regression function $f_0: \Xset \to \Reals$. The user then builds a neighborhood graph over the design points, so that ${\bf W}_{ij}$ is large iff $X_i$ and $X_j$ are close, say in Euclidean distance. In this context, various graph-based estimators have been shown to be \emph{consistent}: as $n \to \infty$, the estimator converges to $f_0$ in an appropriate norm, assuming $f_0$ satisfies some notion of smoothness such as possessing a bounded derivative. However, until recently such statements were not accompanied by error rates, and such error rates as have been proved are not optimal.
\end{itemize}
Viewed as a whole, this body of work leaves a key question unanswered: what are the optimality properties of graph Laplacian smoothing, viewed as an estimator of a (continuum) regression function? This is no small point---rather, it is arguably the fundamental question of nonparametric regression, and without an answer one cannot fully compare the statistical properties of Laplacian smoothing to its competitors.

Our work addresses this gap, and our main contributions can all be summarized as follows: Laplacian smoothing over a random design is minimax optimal over certain first-order \emph{continuum} Sobolev balls. We will show that this is true for the estimator $\wh{f}$ and test statistc $\wh{T}$, when $\Xset \subset \Rd$ is a full-dimensional domain and $1 \leq d \leq 4$. Additionally, we will consider the \emph{manifold assumption}---roughly, that $\Xset$ is a manifold embedded in $\Rd$, with intrinsic dimension $m \ll d$---and derive upper bounds which depend only on the intrinsic dimension $m$, and are independent of the ambient dimension $d$; these bounds too will be optimal. For ease of reading, we provide the upper bounds derived in our main theorems in Table~\textcolor{red}{(?)}.

\paragraph{Organization.}
We now outline the structure of the rest of this paper. In Section~\ref{sec:problem_setup_and_background}, we formalize our model of nonparametric regression with random design, recall the minimax rates for the estimation and goodness-of-fit testing over Sobolev spaces, and review some related work. In Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we give our main results on the optimality of Laplacian smoothing methods, offering some comparison with other nonparametric regression approaches as well as giving insight into our proof techniques. In Section~\ref{sec:simulations} we support our theory with some experiments, before concluding in Section~\ref{sec:discussion}.

\textcolor{red}{(TODO)}:

\section{PROBLEM SETUP AND BACKGROUND}
\label{sec:problem_setup_and_background}
Formally speaking, In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Reals^d$, and 
\begin{equation}
\label{eqn:random_design_regression}
Y_i = f_0(X_i) + \varepsilon_i
\end{equation}
with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise. We will assume the following throughout:
\begin{enumerate}[label=(P\arabic*)]
	\item
	\label{asmp:bounded_lipschitz_density} 
	$P$ admits a density $p$ bounded away from $0$ and $\infty$, i.e.
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty,~~\textrm{for all $x \in \Xset$.}
	\end{equation*}
	Additionally, $p$ is Lipschitz, with Lipschitz constant $L_p$.
\end{enumerate}

\paragraph{Neighborhood graph Laplacians.}
As mentioned, in the graph-based approach to nonparametric regression, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense. The neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$. The $n \times n$ weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $\mathbf{W}_{ij}$ are given by
\begin{equation*}
\label{eqn:neighborhood_graph}
{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}}{r}\Biggr).
\end{equation*}
where $\norm{\cdot} = \norm{\cdot}_{\Rd}$ is the Euclidean norm. Then the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$, and the graph Laplacian can be written as
\begin{equation}
\label{eqn:graph_Laplacian}
\Lap_{n,r} = \bf{D} - \bf{W};
\end{equation}
we add the subscripts $n$ and $r$ to $\Lap$ and $G$ to emphasize that they depend on the random design points $\bf{X}$ and the radius $r$.

\paragraph{Minimax rates over Sobolev spaces.}
Now that our estimator $\wh{f}$ and test statistic $\wh{T}$ are defined in the random design setup, we step away from these methods---indeed, away from graphs entirely--- for a moment, to briefly recall some very classical results regarding minimax rates over Sobolev classes.

\paragraph{Previous work.}

\textcolor{red}{(TODO)}

\section{MINIMAX OPTIMALITY OF LAPLACIAN SMOOTHING}
\label{sec:minimax_optimal_laplacian_smoothing}

\textcolor{red}{(TODO)}: Section 4 of the previous draft.

\paragraph{Comparison with the continuous-time approach.}

\textcolor{red}{(TODO)}: Take some of section 4, and a little of section 2, to do this paragraph.

\paragraph{Overview of analysis.}
\begin{itemize}
	\item Speaking in gross generality, there are two ways in which one can analyze nonparametric regression methods. 
	\item Way (1) --- by far the more common approach to analyzing estimators defined as solutions to variational problems---is to use \emph{empirical process theory} as the workhorse. In other words, one bounds the supremum of the empirical process, i.e.
	\begin{equation}
	\label{eqn:empirical_process}
	\sup_{f \in H^1(\Xset)} \abs{\langle \varepsilon, f \rangle} \lesssim \delta_n
	\end{equation}
	and then uses~\eqref{eqn:empirical_process} to guarantee that the estimator in question will not overfit to the noise $\varepsilon$. While this can give the right rate when $d = 1$, it will totally fail when $d \geq 2$. 
	\item Way (2) is the bias-variance tradeoff. Typically this is not tractable for an estimators such as $\wh{f}_{\LS}$ that is defined as the solution to a variational problem. However the problem~\textcolor{red}{(?)} is special: it has a nice closed-form solution that is amenable to a bias-variance decomposition. We therefore do not need to rely on empirical process theory, and so we will be able to analyze $\wh{f}_{\LS}$ and $T_{\LS}$ beyond the univariate setting.
\end{itemize}

\section{SIMULATIONS}
\label{sec:simulations}

\section{DISCUSSION}
\label{sec:discussion}

\bibliographystyle{plainnat}
\bibliography{../../../graph_testing_bibliography}

\end{document}

### 2020-10-09 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing3.tex
Completed second draft of AISTATS paper on Laplacian smoothing.

@@
 \usepackage{aistats2021}
-\usepackage{amsmath}
-\usepackage{amsfonts, amsthm, amssymb}
+\usepackage{amsmath, amsfonts, amsthm, amssymb}
 \usepackage{graphicx}
@@
 \usepackage{tikz}
+\usepackage{verbatim}
 
@@
 \DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}
+
+% Make a widecheck symbol (thanks, Stack Exchange!)
+\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
+\DeclareFontShape{U}{mathx}{m}{n}{
+	<5> <6> <7> <8> <9> <10>
+	<10.95> <12> <14.4> <17.28> <20.74> <24.88>
+	mathx10
+}{}
+\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
+\DeclareFontSubstitution{U}{mathx}{m}{n}
+\DeclareMathAccent{\widecheck}{0}{mathx}{"71}
+% widecheck made
 
@@
 	In this paper we study the statistical properties of Laplacian smoothing, a graph-based approach to non-parametric regression.
-We show that under standard regularity conditions, the Laplacian smoothing estimator $\wh{f}$, and a test using the statistic $T = \frac{1}{n}\|\wh{f}\|_2^2$, attain the minimax optimal estimation and testing rates of convergence over the first-order Sobolev class $H^1(\Xset)$, where $\Xset \subset \Rd$ and $1 \leq d \leq 4$.
+We show that under standard regularity conditions, the Laplacian smoothing estimator $\wh{f}$, and a test using the statistic $\wh{T} = \frac{1}{n}\|\wh{f}\|_2^2$, attain the minimax optimal estimation and testing rates of convergence over the first-order Sobolev class $H^1(\Xset)$, where $\Xset \subset \Rd$ and $1 \leq d \leq 4$.
 Additionally, we prove that Laplacian smoothing is manifold adaptive: if $\Xset$ is an $m$ dimensional manifold embedded in $\Reals^d$ with $m \ll d$, then the error rate of Laplacian smoothing methods depends only on the intrinsic dimension $m$ and not on the ambient dimension $d$.
@@
 The Laplacian smoothing estimator $\wh{f}$ \citep{smola2003} is a penalized least squares estimator, defined on a graph.
-Letting $G = \bigl(\{1,\ldots,n\},\mathbf{W}\bigr)$ be a weighted graph over the vertices $\{1,\ldots,n\}$,
+Letting $G = \bigl(\{1,\ldots,n\},W\bigr)$ be a weighted graph over the vertices $\{1,\ldots,n\}$,
 \begin{equation}
@@
 \begin{equation*}
-f^T \Lap f = \frac{1}{2} \sum_{i,j = 1}^{n} {\bf W}_{ij}\bigl(Y_i - Y_j\bigr)^2
+f^T \Lap f = \frac{1}{2} \sum_{i,j = 1}^{n} W_{ij}\bigl(Y_i - Y_j\bigr)^2
 \end{equation*}
-encourages $\wh{f}_i \approx \wh{f}_j$ when ${\bf W}_{ij}$ is large. Assuming~\eqref{eqn:laplacian_smoothing} is a reasonable estimator of $f_0$, the squared empirical norm
+encourages $\wh{f}_i \approx \wh{f}_j$ when $W_{ij}$ is large. Assuming~\eqref{eqn:laplacian_smoothing} is a reasonable estimator of $f_0$, the squared empirical norm
 \begin{equation}
@@
 The estimate $\wh{f}$ can be computed by solving a symmetric, diagonally dominant linear system.
-There are known extremely fast and reliable solvers for exactly this problem \textcolor{red}{(TODO)}: references.
-	\item \emph{Generalizability.} Laplacian smoothing is well defined whenever one can associate a graph with some observed responses.
-This generality lends itself to many different data modalities: for instance sentiment analysis and image classification, to name only two.
-\textcolor{red}{(TODO)}: references
-	\item \emph{Supervision.} Although we study Laplacian smoothing in the supervised learning setup~\eqref{eqn:signal_plus_noise_model}, it can be easily adapted to the semi-supervised or unsupervised settings.
+There are known extremely fast and reliable solvers for exactly this problem (see e.g.
+the seminal papers by \cite{spielman2011,spielman2013,spielman2014}, or the review article~\cite{vishnoi2012} and references therein).
+	\item \emph{Generality.} Laplacian smoothing is well defined whenever one can associate a graph with some observed responses.
+This generality lends itself to many different data modalities: for instance text and image classification, to name only two \citep{kondor2002, belkin03a,belkin2006} \textcolor{red}{(TODO)}: Add a bunch more references here.
+	\item \emph{Supervision.} Although we study Laplacian smoothing in the supervised learning setup~\eqref{eqn:signal_plus_noise_model}, it can be easily adapted to the semi-supervised or unsupervised settings \citep{nadler09,slepcev17,dunlop2020,calder2019b}.
 \end{itemize}
@@
 	\item In the \emph{discrete (fixed design) perspective}, both $G$ and $f_0 \in \Reals^n$ are treated as fixed, and the regression ``function'' is really a vector.
-In this context, tight upper bounds have been derived on the error of graph-based estimators, which certify that such estimators are \emph{optimal} over ``function'' classes $\mc{F} \subseteq \Reals^n$.
-The downside of this work is that, depending on the context, it may be unnatural to assume a priori that the graph $G$ is fixed, and that the signal $f_0$ belongs to some discrete ``function'' class.
-	\item In the arguably more natural \emph{continuum (random design) perspective}, in addition to the responses $Y_i$ one observes randomly sampled design points $\{X_1,\ldots,X_n\}$ belonging to a domain $\Xset \subseteq \Rd$, and assumes that $f_{0,i} \equiv f_0(X_i)$ are $n$ evaluations of a regression function $f_0: \Xset \to \Reals$.
-The user then builds a neighborhood graph over the design points, so that ${\bf W}_{ij}$ is large iff $X_i$ and $X_j$ are close, say in Euclidean distance.
-In this context, various graph-based estimators have been shown to be \emph{consistent}: as $n \to \infty$, the estimator converges to $f_0$ in an appropriate norm, assuming $f_0$ satisfies some notion of smoothness such as possessing a bounded derivative.
-However, until recently such statements were not accompanied by error rates, and such error rates as have been proved are not optimal.
+In this context, tight upper bounds have been derived on the error of graph-based estimators \citep{wang2016, sadhanala16,sadhanala17,kirichenko2017,kirichenko2018} and tests \citep{sharpnack2013,sharpnack2013b,sharpnack2015}, which certify that such procedures are \emph{optimal} over ``function'' classes $\mc{F} \subseteq \Reals^n$.
+The downside of this work is that, depending on the context, it may be unnatural to assume a priori that the graph $G$ is fixed, or that the signal $f_0$ belongs to some discrete ``function'' class.
+	\item In the arguably more natural \emph{continuum (random design) perspective}, in addition to the noisy responses $Y_i$ one observes randomly sampled design points $\{X_1,\ldots,X_n\}$ belonging to a domain $\Xset \subseteq \Rd$, and assumes that $f_{0,i} \equiv f_0(X_i)$ are $n$ evaluations of a regression function $f_0: \Xset \to \Reals$.
+The user then builds a neighborhood graph over the design points, so that $W_{ij}$ is large iff $X_i$ and $X_j$ are close, say in Euclidean distance.
+In this context, various graph-based procedures have been shown to be \emph{consistent}: as $n \to \infty$, they converge to a \emph{continuum} limit (see \citep{belkin07,trillos2018,vonluxburg2008} among others).
+However, until recently such statements were not accompanied by error rates, and such error rates as have been proved \citep{lee2016,trillos2020} are not optimal.
 \end{itemize}
@@
 Our work addresses this gap, and our main contributions can all be summarized as follows: Laplacian smoothing over a random design is minimax optimal over certain first-order \emph{continuum} Sobolev balls.
-We will show that this is true for the estimator $\wh{f}$ and test statistc $\wh{T}$, when $\Xset \subset \Rd$ is a full-dimensional domain and $1 \leq d \leq 4$.
-Additionally, we will consider the \emph{manifold assumption}---roughly, that $\Xset$ is a manifold embedded in $\Rd$, with intrinsic dimension $m \ll d$---and derive upper bounds which depend only on the intrinsic dimension $m$, and are independent of the ambient dimension $d$; these bounds too will be optimal.
-For ease of reading, we provide the upper bounds derived in our main theorems in Table~\textcolor{red}{(?)}.
+We will show that this is true for the estimator $\wh{f}$ and test statistic $\wh{T}$, when $\Xset \subset \Rd$ is a full-dimensional domain and $1 \leq d \leq 4$.
+Additionally, we will consider the \emph{manifold hypothesis}---roughly, that $\Xset$ is a manifold embedded in $\Rd$, with intrinsic dimension $m \ll d$---and derive upper bounds which depend only on the intrinsic dimension $m$, and are independent of the ambient dimension $d$; these bounds too will be optimal.
+For ease of reading, we provide the upper bounds derived in all our main theorems in Table~\textcolor{red}{(1)}.
+
+\textcolor{red}{(TODO)}: Make Table 1.
 
@@
 We now outline the structure of the rest of this paper.
-In Section~\ref{sec:problem_setup_and_background}, we formalize our model of nonparametric regression with random design, recall the minimax rates for the estimation and goodness-of-fit testing over Sobolev spaces, and review some related work.
-In Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we give our main results on the optimality of Laplacian smoothing methods, offering some comparison with other nonparametric regression approaches as well as giving insight into our proof techniques.
-In Section~\ref{sec:simulations} we support our theory with some experiments, before concluding in Section~\ref{sec:discussion}.
+In Section~\ref{sec:problem_setup_and_background}, we formalize our model of nonparametric regression with random design, and recall the minimax rates for estimation and goodness-of-fit testing over Sobolev spaces.
+In Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we give our main results on the optimality of Laplacian smoothing methods, then draw some connections to other nonparametric regression approaches and summarize our proof techniques.
+In Section~\ref{sec:manifold_adaptivity} we show that Laplacian smoothing optimally leverages the manifold hypothesis.
+Then in Section~\ref{sec:simulations} we support our theory with some experiments, before concluding in Section~\ref{sec:discussion}.
 
-\textcolor{red}{(TODO)}:
+\paragraph{Notation.}
+The graph Laplacian $L$ is a symmetric, positive semi-definite matrix, and admits an eigendecomposition $L = \sum_{k = 1}^{n} \lambda_k v_k v_k^T$.
+We order the eigenvalues $0 = \lambda_1 \leq \lambda_2 \leq \cdots \leq \lambda_n$, and always consider unit-norm eigenvectors $\|v_k\|_2^2 = 1$; we will make reference to these objects later on.
+For a function $f: \Xset \to \Reals^n$ and design points $X_1,\ldots,X_n$, the empirical norm is $\|f\|_n^2 := \frac{1}{n}\sum_{i = 1}^{n} f(X_i)^2$.
+For an integer $p > 0$, we use $\Leb^p(\Xset)$ to refer to the set of functions $f$ for which the $p$th moment $\norm{f}_{\Leb^p(\Xset)}^p := \int_{\Xset} |f(x)|^p \,dx < \infty$, and $C^p(\Xset)$ to refer to those function which are $p$-times continuously differentiable with bounded $p$th derivative.
+Finally, we use $C,C_1,\ldots$ and $c,c_1,\ldots$ as constants which are independent of $n$ and $f_0$, and for convenience, we will also use asymptotic notation: $a_n \lesssim b_n$ means that $a_n \leq Cb_n$ for some $C > 0$, and $a_n \asymp b_n$ means that $a_n \lesssim b_n$ and $b_n \lesssim a_n$.
 
@@
 \label{sec:problem_setup_and_background}
-Formally speaking, In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Reals^d$, and
+Formally speaking, in the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Reals^d$, and
 \begin{equation}
@@
 \end{equation}
-with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise. We will assume the following throughout:
+with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise.
+
+\paragraph{Neighborhood graph Laplacians.}
+As mentioned, in the graph-based approach to nonparametric regression we first build a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense.
+The neighborhood graph $G_{n,r} = ([n],W)$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$.
+The $n \times n$ weight matrix $W = (W_{ij})$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $W_{ij}$ are given by
+\begin{equation*}
+\label{eqn:neighborhood_graph}
+W_{ij} = K\Biggl(\frac{\norm{X_i - X_j}}{r}\Biggr).
+\end{equation*}
+where $\norm{\cdot} = \norm{\cdot}_{\Rd}$ is the Euclidean norm.
+Then the degree matrix $D$ is the $n \times n$ diagonal matrix with entries $D_{ii} = \sum_{j = 1}^{n}W_{ij}$, and the graph Laplacian can be written as
+\begin{equation}
+\label{eqn:graph_Laplacian}
+\Lap_{n,r} = D - W;
+\end{equation}
+we add the subscripts $n$ and $r$ to $\Lap$ and $G$ to emphasize that they depend on the random design points $X_1,\ldots,X_n$ and the radius $r$.
+To avoid any confusion: henceforth when speak of $\wh{f}$ and $\wh{T}$, we will always assume they are fit using the neighborhood graph Laplacian $L = L_{n,r}$.
+
+\paragraph{Sobolev spaces.}
+Now that we've defined the neighborhood graph Laplacian $\Lap_{n,r}$, we step away from graph-based methods for a moment, to briefly recall some very classical results regarding minimax rates over Sobolev classes.
+
+We say that a function $f \in \Leb^2(\Xset)$ belongs to the \emph{first-order Sobolev space} $H^1(\Xset)$ iff for each $j = 1,\ldots,d$, the weak partial derivative $D^jf$ exists and belong to $\Leb^2(\Xset)$.
+For such functions $f \in H^1(\Xset)$, the Sobolev seminorm $\seminorm{f}_{H^{1}(\Xset)}$ is the average size of the gradient;
+\begin{equation*}
+\seminorm{f}_{H^1(\Xset)}^2 = \int_{\Xset} \bigl\|\nabla f(x)\bigr\|^2 \,dx
+\end{equation*}
+with corresponding Sobolev norm $\norm{f}_{H^1(\Xset)} := \norm{f}_{\Leb^2(\Xset)} + |f|_{H^1(\Xset)}$. The Sobolev balls $H^1(\Xset,M)$ for $M > 0$ are then defined as
+\begin{equation*}
+H^1(\Xset,M) := \Bigl\{f \in H^1(\Xset): \norm{f}_{H^1(\Xset)}^2 \leq M^2\Bigr\}.
+\end{equation*}
+
+\paragraph{Minimax rates.}
+To carry out a minimax analysis of Sobolev spaces, one must impose some regularity conditions on the design distribution $P$. 
 \begin{enumerate}[label=(P\arabic*)]
+	\item
+	\label{asmp:domain}
+	$P$ is supported on a domain $\Xset \subset \Rd$, an open, connected set with Lipschitz boundary.
 	\item
@@
 	\end{equation*}
-	Additionally, $p$ is Lipschitz, with Lipschitz constant $L_p$.
+	Additionally, the restriction of $p$ to $\Xset$ is Lipschitz, with Lipschitz constant $L_p$.
 \end{enumerate}
 
-\paragraph{Neighborhood graph Laplacians.}
-As mentioned, in the graph-based approach to nonparametric regression, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense.
-The neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$.
-The $n \times n$ weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $\mathbf{W}_{ij}$ are given by
+Under~\ref{asmp:bounded_lipschitz_density} and~\ref{asmp:domain}, the minimax estimation rate over Sobolev balls is (see e.g. \citep{tsybakov2008_book})
+\begin{equation}
+\label{eqn:sobolev_space_estimation_minimax_rate}
+\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{d/(2 + d)}n^{-2/(2 + d)}
+\end{equation}
+
+As nonparametric hypothesis testing is (comparatively) less familiar than nonparametric estimation, we briefly summarize the main idea before stating the optimal error rate.
+In the goodness-of-fit testing problem, we ask for a test function which can distinguish between the null and alternative hypotheses
+\begin{equation}
+\mathbf{H}_0: f_0 = 0, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{0\}
+\end{equation} 
+where $\mc{F}$ is some class of functions.
+The minimax critical radius is the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{F}_{\epsilon} := \mc{F} \cap \{f: \|f\|_{\Leb^2(\Xset)} \geq \epsilon\}$:
 \begin{equation*}
-\label{eqn:neighborhood_graph}
-{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}}{r}\Biggr).
+\epsilon(\mc{F}) := \inf\Biggl\{\epsilon > 0: \inf_{\phi} \biggl[\Ebb_0[\phi] +  \sup_{f_0 \in \mc{F}} \Ebb_{f_0}[1 - \phi]\biggr] \leq 2\alpha\Biggr\}
 \end{equation*} 
-where $\norm{\cdot} = \norm{\cdot}_{\Rd}$ is the Euclidean norm.
-Then the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$, and the graph Laplacian can be written as
+where $\Ebb_{f_0}[\cdot]$ is the expectation under the regression function $f_0$. See \cite{ariascastro2018} for a more extended discussion of the minimax paradigm in nonparametric testing.
+
+Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing critical radius over $H^1(\Xset,M)$ is much smaller than the minimax estimation rate \citep{ingster2009}:
 \begin{equation}
-\label{eqn:graph_Laplacian}
-\Lap_{n,r} = \bf{D} - \bf{W};
+\label{eqn:sobolev_space_testing_critical_radius}
+\epsilon^2\Bigl(H^1(\Xset,M)\Bigr) \asymp M^{2d/(4 + d)}n^{-4/(4 + d)}~~\textrm{for $1 \leq d < 4$.}
 \end{equation}
-we add the subscripts $n$ and $r$ to $\Lap$ and $G$ to emphasize that they depend on the random design points $\bf{X}$ and the radius $r$.
-
-\paragraph{Minimax rates over Sobolev spaces.}
-Now that our estimator $\wh{f}$ and test statistic $\wh{T}$ are defined in the random design setup, we step away from these methods---indeed, away from graphs entirely--- for a moment, to briefly recall some very classical results regarding minimax rates over Sobolev classes.
-
-\paragraph{Previous work.}
-
-\textcolor{red}{(TODO)}
+When $d \geq 4$ the functions in $H^1(\Xset)$ are very irregular---formally speaking $H^1(\Xset)$ does not continuously embed into $\Leb^4(\Xset)$ when $d \geq 4$---and the minimax rates in this regime are unknown.
 
@@
 
-\textcolor{red}{(TODO)}: Section 4 of the previous draft.
+We now formalize the main conclusions of this work: that Laplacian smoothing methods on neighborhood graphs are minimax rate-optimal over first-order continuum Sobolev classes.
+Throughout this section, we will assume~\ref{asmp:bounded_lipschitz_density} and~\ref{asmp:domain}.
+Additionally, we will assume the kernel $K$ satisfies condition~\ref{asmp:kernel}.
+\begin{enumerate}[label=(K\arabic*)]
+	\item
+	\label{asmp:kernel}
+	$K:[0,\infty) \to [0,\infty)$ is a non-increasing function supported on $[0,1]$, its restriction to $[0,1]$ is Lipschitz, and $K(1) > 0$. Additionally, it is normalized so that
+	\begin{equation*}
+	\int_{\Reals^d} K\bigl(\norm{z}\bigr) \,dz = 1.
+	\end{equation*}
+	Let $\sigma_K := \frac{1}{d} \int_{\Rd} \|x\|^2 K(\|x\|) \,dx$.
+\end{enumerate}
+This is a relatively mild condition: the choice of kernel is under the user's control, and moreover~\ref{asmp:kernel} covers many (though certainly not all) common kernel choices.
 
-\paragraph{Comparison with the continuous-time approach.}
+\paragraph{Estimation error of Laplacian smoothing.} 
+Under these conditions, the Laplacian smoothing estimator $\wh{f}$ achieves, up to constants, the minimax risk $n^{-2/(2 + d)}$ over all $f_0 \in H^1(\Xset,M)$.
+This statement will hold whenever the graph $G_{n,r}$ is computed with radius $r$ in the following range.
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{0}
+	\item 
+	\label{asmp:ls_kernel_radius_estimation}
+	The neighborhood graph radius $r$ satisfies
+	\begin{equation*}
+	\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq r \leq n^{-3/(4 + 2d)} M^{(d - 4)/(4 + 2d)} \wedge c_0.
+	\end{equation*}
+\end{enumerate}
+We comment on~\ref{asmp:ls_kernel_radius_estimation} after stating Theorem~\ref{thm:laplacian_smoothing_estimation1}.
+\begin{theorem}
+	\label{thm:laplacian_smoothing_estimation1}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d < 4$ and $f_0 \in H^1(\Xset,M)$.
+Suppose that that the neighborhood graph $G_{n,r}$ is computed with a radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_estimation},  and the Laplacian smoothing estimator $\wh{f}$ with $\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$.
+It holds that
+	\begin{equation*}
+	\Bigl\|\wh{f} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{d/(2 + d)} n^{-2/(2 + d)}
+	\end{equation*}
+	with probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c M^{d/(2d + 4)} n^{d/(2+d)})$.
+\end{theorem}
+To summarize: when $d = 1,2$ or $3$, the Laplacian smoothing estimator $\wh{f}$ has in-sample mean squared error within a constant factor of the minimax risk, except on a set of small probability.
+Some remarks:
+\begin{itemize}
+	\item When $d = 4$, by setting $r \asymp (\log n/n)^{1/4}$ our analysis results in an upper bound on the mean squared error within a $\log n$ factor of the minimax risk, but when $d \geq 5$ our analysis yields upper bounds that are much worse than the minimax rate.
+This mirrors the conclusions of~\cite{sadhanala16}, who investigate estimation rates of Laplacian smoothing over the $d$-dimensional grid graph.
+\cite{sadhanala16} argue that their analysis is tight, and that it is the estimator, rather than the analysis, which is deficient when $d \geq 5$.
+Formally proving such a claim turns out to be harder in the random design setting than in the fixed design setting; although we suspect a similar claim is true, we cannot prove it, and instead investigate the matter empirically in Section~\ref{sec:simulations}.
+	\item The lower bound on $r$ imposed by~\ref{asmp:ls_kernel_radius_estimation} is  compatible with practice, where by far the most common choice of radius is the connectivity threshold $r \asymp (\log(n)/n)^{1/d}$, chosen to make the graph $G_{n,r}$ as sparse as possible while still being connected, for maximum computational efficiency.
+The upper bound is a bit more mysterious.
+We need it for technical reasons to ensure that $\wh{f}$ does not overfit, but note that as a practical matter one rarely chooses $r$ to be so large.
+	\item Since $\wh{f}$ is defined only at the design points $X_1,\ldots,X_n$, we measure loss using the in-sample mean squared error $\norm{\cdot}_n^2$.
+Of course the in-sample error is a random variable, and our bound is thus a bound in high probability.
+That being said, it is possible to smoothly extend $\wh{f}$ to be defined on all of $\Xset$, or indeed all of $\Reals^d$---for instance, using the \emph{Nystrom extension}---and evaluate the error of the extension in $\Leb^2(\Xset)$ norm.
+Assuming that the extension of our estimators and the regression function $f_0$ are suitably smooth, tools from empirical process theory should guarantee that the $\Leb^2(\Xset)$ error is not too much greater than the in-sample error, but we do not further pursue the details here.
+\end{itemize}
 
-\textcolor{red}{(TODO)}: Take some of section 4, and a little of section 2, to do this paragraph.
+\paragraph{Testing error of Laplacian smoothing.}
+Let us define a test using the statistic $\wh{T}$. For $b \geq 1$, set the threshold $\wh{t}_b$ as
+\begin{equation*}
+\wh{t}_{b} := \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k + 1\bigr)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k + 1\bigr)^4}},
+\end{equation*}
+where we recall $\lambda_k$ is the $k$th smallest eigenvalue of $\Lap_{n,r}$; the Laplacian smoothing test is simply $\wh{\varphi} := \1\bigl\{\wh{T} \leq \wh{t}_b\bigr\}$.
+(Clearly the test $\wh{\varphi}$ depends on $b$, but we suppress this notationally.) The number $b \geq 1$ is a tuning parameter selected by the user, with the choice made based on the tolerated level of Type I and Type II error.
+In fact, the threshold $\wh{t}_b$ is precisely the right choice to control the Type I error of $\wh{\varphi}$; as we show in Lemma~\ref{lem:ls_fixed_graph_testing} in the supplementary material,
+\begin{equation*}
+\Ebb_0\bigl[\wh{\varphi}\bigr] \leq \frac{1}{b^2}.
+\end{equation*}
+Theorem~\ref{thm:laplacian_smoothing_testing} establishes that this test also has small Type II error, uniformly over all $f_0$ separated from $0$ by at least the critical radius $\epsilon\bigl(H^1(\Xset,M)\bigr)$ given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
+For this to hold, we will require a tight range of scalings for $r$.
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{1}
+	\item 
+	\label{asmp:ls_kernel_radius_testing}
+	The neighborhood graph radius $r$ satisfies
+	\begin{equation*}
+	\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq r \leq M^{(16 - d)/(8 + 2d)}n^{(d - 20)/(32 + 8d)} \wedge c_0.
+	\end{equation*}
+\end{enumerate}
+
+\begin{theorem}
+	\label{thm:laplacian_smoothing_testing}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d < 4$ and $f_0 \in H^1(\Xset,M)$ for $M \leq n^{(4 - d)/(4 + d)}$.
+Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}, and the Laplacian smoothing test $\wh{\varphi}$ with $\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$, and $b \geq 1$.
+For all $f_0$ such that
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_testing}
+	\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2d/(4 + d)} n^{-4/(4 + d)}
+	\end{equation} 
+	the Type II error is upper bounded:
+	\begin{align*}
+	& \Ebb_{f_0}\Bigl[1 - \wh{\varphi} \Bigr] \leq \\
+	& C\biggl(\frac{1}{b}\Bigl[1 + n^{-d(4 + d)}M^{-2d/(4 + d)}\Bigr] + C_1n\exp\bigl(-c_1nr^d\bigr)\biggr).
+	\end{align*}
+\end{theorem}
+A couple remarks:
+\begin{itemize}
+	\item As mentioned in Section~\ref{sec:problem_setup_and_background}, when $d \geq 4$ the Sobolev balls $H^1(\Xset,M)$ include quite irregular functions $f \not\in \Leb^4(\Xset)$.
+Proving tight lower bounds over such classes is non-trivial, and to the best of our knowledge such an analysis remains outstanding.
+On the other hand, if we explicitly assume that $f_0 \in \Leb^4(\Xset,M)$, then \cite{guerre02} show that the testing problem is characterized by the dimension-free lower bound $\epsilon^{2}(\Leb^4(\Xset,M)) \gtrsim n^{-1/2}$.
+Moreover, by training $\wh{f}$ to the interpolation limit, i.e.
+setting $\rho = 0$, the test $\wh{\varphi}$ achieves (up to constants) this lower bound.
+That is, for any $f_0 \in \Leb^4(\Xset,M)$ such that $\|f_0\|_{\Leb^2(\Xset)}^2 \geq C b^2n^{-1/2}$, we have
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_testing_low_smoothness}
+	E_{f_0}\Bigl[1 - \wh{\varphi}\Bigr] \leq \frac{C(1 + M^4)}{b^2}
+	\end{equation} 
+	\item To compute the data-dependent threshold $\wh{t}_b$, one must know each of the eigenvalues $\lambda_1,\ldots,\lambda_n$.
+Computing all $n$ of these eigenvalues is far more expensive---on the order of $n^3$ operations---than computing $\wh{T}$.
+That being said, in practice we would not recommend using $\wh{t}_b$ anyway: the threshold is justified by asymptotic theory, but in practice $n$ is always finite, and using such thresholds can result in substantial loss of efficiency.
+We would instead make the standard recommendation to calibrate via permutation.
+\end{itemize}
+
+\paragraph{Connections with variational problems.}
+We now connect Laplacian smoothing to some related approaches for nonparametric regression, and in so doing reveal some surprising aspects of Theorems~\ref{thm:laplacian_smoothing_estimation1} and~\ref{thm:laplacian_smoothing_testing}.
+To keep things simple, we will focus on the estimation context, but everything we say will apply to testing as well.
+
+Note that as an alternative to Laplacian smoothing --- which encourages $\wh{f}_1,\ldots,\wh{f}_n$ to satisfy a notion of smoothness over $G_{n,r}$--- we could have directly asked for an estimate $\wt{f}$ which was smooth in the Sobolev sense,
+\begin{equation}
+\label{eqn:thin_plate_spline}
+\wt{f} := \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \cdot \seminorm{f}_{H^{1}(\Xset)}^2\biggr\}
+\end{equation}
+where the minimum is taken over all $f \in H^1(\Xset)$.
+Connections between the penalties in~\eqref{eqn:laplacian_smoothing} and~\eqref{eqn:thin_plate_spline} have been known for some time: \cite{bousquet03} show that for any $f \in C^2(\Xset)$,
+\begin{equation}
+\label{eqn:seminorm_consistency}
+\begin{aligned}
+\lim \frac{1}{n^2 r^{d + 2}} f^T \Lap_{n,r} f & = \int_{\Xset} f(x) \cdot \Delta_Pf(x) p(x) \,dx \\
+& = \int_{\Xset} \|\nabla f(x)\|^2 p^2(x) \,dx \\
+& = |f|_{H^1(\Xset)}^2;
+\end{aligned}
+\end{equation}
+in the above, the limit is as $n \to \infty$ and $r \to 0$, $\Delta_P: C^2(\Xset) \to \Leb^2(\Xset)$ is the Laplace-Beltrami operator
+\begin{equation*}
+\Delta_pf(x) := -\frac{1}{p} \dive\bigl(p^2\nabla f(x))
+\end{equation*}
+and the second equality follows if $f$ satisfies Dirichlet boundary conditions, by integrating by parts.
+
+Equation~\eqref{eqn:seminorm_consistency} justifies thinking of $\wh{f}$ as a noisy, discrete approximation to $\wt{f}$.
+At first blush, the penalty $|f|_{H^1(\Xset)}^2$ seems to leverage the assumption that $f_0 \in H^1(\Xset,M)$ more directly than the penalty $f^T \Lap_{n,r} f$, and so the estimator $\wt{f}$ appears a more natural candidate than $\wh{f}$ for regression over Sobolev spaces.
+We might therefore expect the benefit of using $\wh{f}$ as opposed to $\wt{f}$ to lie purely in its computational properties, and hope only that in using $\wh{f}$ as opposed to $\wt{f}$ we are not forced to sacrifice statistical efficiency.
+
+In reality, however, the story is actually quite different: remarkably, $\wh{f}$ exhibits \emph{superior} statistical properties to $\wt{f}$.
+Indeed, the estimator $\wt{f}$ is extremely well known--- in the univariate case $d = 1$ it is simply a \emph{smoothing spline}, when $d > 1$ it is an example of a \emph{thin-plate spline}---and its statistical properties are by this point well understood.
+When $d = 1$, the error of $\wt{f}$ is upper bounded $\|\wt{f} - f_0\|_n^2 \lesssim n^{-2/3}$ with high probability, and the estimator is minimax rate optimal (see e.g.
+\cite{vandergeer2000}, or \cite{liu2019} for the analogous statement w.r.t testing).
+However, when $d \geq 2$, $\wt{f}$ is not merely suboptimal in the minimax sense, but not even well-posed.
+It is known (see e.g.
+\cite{green93}) that by placing ``bump functions'' of arbitrarily small radius around each of the design points $X_1,\ldots,X_n$, one can interpolate the responses $Y_1,\ldots,Y_n$ in such a manner that forces the criterion in~\eqref{eqn:thin_plate_spline} to 0; such an estimator, of course, will be inconsistent.
+Therefore when $d = 2,3$ or $4$, we see that Laplacian smoothing attains minimax optimal rates\footnote{Up to a factor of $\log n$ when $d = 4$.} in a setting where its continuum counterpart, thin-plate splines, are not even well-defined.
+
+What is going on?
+The key point is that the discretization imposed by the graph $G_{n,r}$ turns out to be a blessing.
+The problem with~\ref{eqn:thin_plate_spline} is that the function class $H^1(\Xset)$ is far too large when $d \geq 2$; we know from the Sobolev embedding theorem \cite{evans10} that it does not continuously embed into any Holder space, and that it is not an Reproducing Kernel Hilbert Space.
+On the other hand~\eqref{eqn:laplacian_smoothing} is a finite dimensional (if growing linearly with $n$) problem---as a result $\wh{f}$ has far less capacity to overfit than does $\wt{f}$.
+Of course, discretization is not the only way to make the problem~\eqref{eqn:thin_plate_spline} more tractable: for instance, one could replace the penalty $|\cdot|_{H^1(\Xset)}$ by the stricter (semi)-norm $|\cdot|_{C^1(\Xset)}$, or conduct the problem over some finite dimensional linear subspace of $H^1(\Xset)$ (i.e use a sieve).
+These solutions improve the statistical properties of $\wt{f}$ when $d > 1$ \citep{birge1993,birge1998,vandergeer2000}, but in practice it is not at all clear how to fit them, in stark contrast to $\wh{f}$.
+
+To be clear, the story is not entirely favorable to Laplacian smoothing.
+The problem~\eqref{eqn:thin_plate_spline} is only a special, first-order case of thin-plate splines.
+In general, the degree-$k$ thin plate spline is defined as a minimizer of
+\begin{equation*}
+\sum_{i = 1}^{n} (Y_i - f(X_i))^2 + \rho \cdot |f|_{H^k(\Xset)}^2
+\end{equation*}
+where $|f|_{H^k(\Xset)}^2$ is the order-$k$ Sobolev seminorm, which measures the size of the order-$k$ partial derivatives in $L^2(\Xset)$ norm.
+Assuming $f_0 \in H^k(\Xset)$ for some $2k > d$, the degree-$k$ thin plate spline has error on the order of $n^{-2k/(2k + d)}$ \cite{vandergeer2000}, which is the minimax optimal rate.
+Of course, assuming $f_0 \in H^k(\Xset)$ for some $2k > d$ is a rather strong assumption which may not be warranted, but at present we do not know whether (adaptations of) Laplacian smoothing on neighborhood graphs can achieve these faster rates.
 
 \paragraph{Overview of analysis.}
-\begin{itemize}
-	\item Speaking in gross generality, there are two ways in which one can analyze nonparametric regression methods. 
-	\item Way (1) --- by far the more common approach to analyzing estimators defined as solutions to variational problems---is to use \emph{empirical process theory} as the workhorse.
-In other words, one bounds the supremum of the empirical process, i.e.
+The previous discussion shows some surprising strengths of the Laplacian smoothing estimator, but it also highlights the difficulty of our analysis task.
+A natural way to analyze $\wh{f}$ would be to use~\eqref{eqn:seminorm_consistency} to establish a coupling with $\wt{f}$, but this will not do the job: first, proving such a coupling is tricky---note that~\eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:thin_plate_spline} is $H^1(\Xset)$---and second, as we've just established we actually would like to bound $\wh{f}$ in settings where we know $\wt{f}$ fails.
+
+Instead we take a different approach, and directly analyze the error of $\wh{f}$ and $\wh{T}$ using a (conditional on $X_1,\ldots,X_n$) bias-variance decomposition.
+That is, we show with high probability that
+\begin{equation*}
+\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \underbrace{\frac{2\rho}{n} \bigl(f_0^T \Lap_{n,r} f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{10}{n} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2}}_{\textrm{variance}}
+\end{equation*}
+and likewise that $\wh{\varphi}$ has high power whenever
+\begin{equation*}
+\bigl\|f_0\bigr\|_n^2 \geq \underbrace{\frac{2\rho}{n} \bigl(f_0^T \Lap_{n,r} f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{4b}{n} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}}_{\textrm{variance}}
+\end{equation*}
+
+Note that the bias and variance terms are each functions of the random graph $G_{n,r}$, and so are themselves random.
+To upper bound them, we build on some recent works \citep{burago2014,trillos2019,calder2019} regarding the consistency of neighborhood graphs to establish the following two Lemmas.
+In the second of these Lemmas, we will assume
+\begin{enumerate}[label=(R\arabic*)]
+	\item
+	\label{asmp:radius1}
+	The radius $r$ satisfies $(\log n/n)^{1/d} \leq r \leq c_0$.
+\end{enumerate}
+and put $A_{n,r}(k) := \min\{nr^{d + 2}k^{2/d},nr^d\}$.
+
+\begin{lemma}
+	\label{lem:graph_sobolev_seminorm}
+	For any $f \in H^1(\Xset)$, 
 	\begin{equation}
-	\label{eqn:empirical_process}
-	\sup_{f \in H^1(\Xset)} \abs{\langle \varepsilon, f \rangle} \lesssim \delta_n
+	\label{eqn:graph_sobolev_seminorm}
+	f^T \Lap_{n,r} f \leq \frac{(1 + L_p)^2 \sigma_K}{\delta} n^2 r^{d + 2} |f|_{H^1(X)}^2
 	\end{equation}
-	and then uses~\eqref{eqn:empirical_process} to guarantee that the estimator in question will not overfit to the noise $\varepsilon$.
-While this can give the right rate when $d = 1$, it will totally fail when $d \geq 2$.
-	\item Way (2) is the bias-variance tradeoff.
-Typically this is not tractable for an estimators such as $\wh{f}_{\LS}$ that is defined as the solution to a variational problem.
-However the problem~\textcolor{red}{(?)} is special: it has a nice closed-form solution that is amenable to a bias-variance decomposition.
-We therefore do not need to rely on empirical process theory, and so we will be able to analyze $\wh{f}_{\LS}$ and $T_{\LS}$ beyond the univariate setting.
-\end{itemize}
+	with probability at least $1 - \delta$.
+\end{lemma}
+\begin{lemma}
+	\label{lem:neighborhood_eigenvalue}
+	For any neighborhood graph radius $r$ which satisfies~\ref{asmp:radius1}, it holds that
+	\begin{equation}
+	\label{eqn:neighborhood_eigenvalue}
+	c_2A_{n,r}(k) \leq \lambda_k \leq C_2A_{n,r}(k)~~\textrm{for all $2 \leq k \leq n$}
+	\end{equation}
+	with probability at least $1 - C_1n\exp(-c_1nr^d)$.
+\end{lemma}
+Lemma~\ref{lem:graph_sobolev_seminorm} gives a direct upper bound on the bias term.
+Lemma~\ref{lem:neighborhood_eigenvalue} results in a sufficient upper bound on the variance term whenever the neighborhood graph radius $r$ is \textit{sufficiently small}; precisely, when $r$ is upper bounded as in~\ref{asmp:ls_kernel_radius_estimation} (for estimation) or~\ref{asmp:ls_kernel_radius_testing} (for testing).
+The parameter $\rho$ is then tuned to minimize the sum of bias and variance, in the usual manner.
+
+\textcolor{red}{(For Siva)}: I think the below paragraph as currently written is not good enough.
+My question(s) for you: do you agree, and if so, do you think it can be fixed by rewriting or should be shelved entirely?
+
+It is useful to give another perspective on our approach.
+In one sense, the bias and variance terms just defined reflect a standard tradeoff commonly seen in the analysis of penalized least squares estimators.
+As the tuning parameter $\rho$ grows, the estimator is better able to approximate the regression function $f_0$, but is more vulnerable to (over)fitting to noise.
+However, when analyzing such estimators one typically assumes two properties: first, that the regression function $f_0$ lies in (or near) a ball in the normed space induced by the penalty, and second that this ball is reasonably small, e.g.
+as measured by metric entropy.
+In contrast, the Laplacian smoothing penalty induces a ball
+\begin{equation*}
+H^1(G_{n,r},M) := \{f: f^T \Lap_{n,r} f \leq M^2\}
+\end{equation*}
+that is data-dependent and random, and so we do not have access to either of the aforementioned properties \emph{a priori}; instead we must prove they hold with high probability.
+In this sense, our analysis is rather unusual.
+
+\section{MANIFOLD ADAPTIVITY}
+\label{sec:manifold_adaptivity}
+The minimax rates $n^{-2/(2 + d)}$ (in estimation) and $n^{-4/(4 + d)}$ (in testing) suffer from the curse of dimensionality.
+One way to overcome this is the \emph{manifold hypothesis}: roughly speaking, that the data $X_1,\ldots,X_n$ lie on a submanifold $\Xset$ of $\Rd$ that has intrinsic dimension $1 \leq m \leq d$.
+Under such an assumption, it is known \citep{bickel2007, ariascastro2018} that the optimal rates over the Lipschitz classes $C^1(\Xset)$ now scale like $n^{-2/(2 + m)}$ (for estimation) and $n^{-4/(4 + m)}$ (for testing), which can be much smaller than the full-dimensional error rates when $m \ll d$.
+
+On the other hand, a theory has been developed~\citep{belkin03,belkin05,belkin2006,niyogi2013} establishing the the neighborhood graph $G_{n,r}$ and Laplacian $\Lap_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+In this section, we contribute to this line of work by showing that Laplacian smoothing on a neighborhood graph achieves the sharper minimax rates over the first-order Sobolev classes $H^1(\Xset)$ under the manifold hypothesis.
+
+\paragraph{Error rates for Laplacian smoothing with the manifold hypothesis.}
+The conditions and results of this Section will be largely similar to the previous one, except with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
+For the rest of this section, we assume the following.
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{2}
+	\item 
+	\label{asmp:domain_manifold}
+	The measure $P$ is supported on a compact, connected, smooth submanifold $\Xset$ of $\Reals^d$.
+The manifold $\Xset$ is of fixed dimension $1 \leq m \leq d$, is without boundary, and has positive reach.
+	\item 
+	\label{asmp:density_manifold} $P$ has a density $p$ with respect to the volume form of $\Xset$ which is bounded away from $0$ and $\infty$,
+	\begin{equation*}
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty
+	\end{equation*}
+	for all $x \in \Xset$. The density $p$ is Lipschitz with Lipschitz constant $L_p$.
+\end{enumerate}
+Under the assumptions~\ref{asmp:domain_manifold} and~\ref{asmp:density_manifold}, and for a suitable range of $r$, the error bounds on estimator $\wh{f}$ and test $\wh{\varphi}$ will depend on $m$ instead of $d$.
+
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{3}
+	\item 
+	\label{asmp:ls_kernel_radius_estimation_manifold}
+	The neighborhood graph radius $r$ satisfies
+	\begin{equation*}
+	\biggl(\frac{\log n}{n}\biggr)^{1/m} \leq r \leq n^{-3/(4 + 2m)} M^{(m - 4)/(4 + 2m)} \wedge c_3.
+	\end{equation*}
+\end{enumerate}
+\begin{theorem}
+	\label{thm:laplacian_smoothing_estimation_manifold}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $m < 4$ and $f_0 \in H^1(\Xset,M)$.
+Suppose that that the neighborhood graph $G_{n,r}$ is computed with a radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_estimation_manifold},  and the Laplacian smoothing estimator $\wh{f}$ with $\rho = M^{-4/(2 + m)} (nr^{m + 2})^{-1} n^{-2/(2 + m)}$.
+It holds that
+	\begin{equation*}
+	\Bigl\|\wh{f} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{m/(2 + m)} n^{-2/(2 + m)}
+	\end{equation*}
+	with probability at least $1 - \delta -  C_4n\exp(-c_4nr^m) - \exp(-c M^{m/(2m + 4)} n^{m/(2+m)})$.
+\end{theorem}
+
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{4}
+	\item 
+	\label{asmp:ls_kernel_radius_testing_manifold}
+	The neighborhood graph radius $r$ satisfies
+	\begin{equation*}
+	\biggl(\frac{\log n}{n}\biggr)^{1/m} \leq r \leq M^{\frac{(16 - m)}{8 + 2m}}n^{\frac{m - 20}{32 + 8m}} \wedge c_3.
+	\end{equation*}
+\end{enumerate}
+
+\begin{theorem}
+	\label{thm:laplacian_smoothing_testing_manifold}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $m < 4$ and $f_0 \in H^1(\Xset,M)$ for $M \leq n^{(4 - m)/(4 + m)}$.
+Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing_manifold}, and the Laplacian smoothing test $\wh{\varphi}$ with $\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$, and $b \geq 1$.
+If
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_testing_manifold}
+	\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2m/(4 + m)} n^{-4/(4 + m)},
+	\end{equation} 
+	the Type II error is upper bounded:
+	\begin{align*}
+	& \Ebb_{f_0}\Bigl[1 - \wh{\varphi} \Bigr] \leq \\
+	& C\biggl(\frac{1}{b}\Bigl[1 + n^{-m(4 + m)}M^{-2m/(4 + m)}\Bigr] + C_4n\exp\bigl(-c_4nr^m\bigr)\biggr).
+	\end{align*}
+\end{theorem}
+The proof of Theorems~\ref{thm:laplacian_smoothing_estimation_manifold} and~\ref{thm:laplacian_smoothing_testing_manifold} proceeds in a similar manner to that of Theorems~\ref{thm:laplacian_smoothing_estimation1} and~\ref{thm:laplacian_smoothing_testing}.
+The key difference is that in the manifold setting, the equations~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} that we use to upper bound bias and variance will hold with $d$ replaced by $m$.
+
+We emphasize that very little about $\Xset$ need be known for Theorems~\ref{thm:laplacian_smoothing_estimation_manifold} and~\ref{thm:laplacian_smoothing_testing_manifold} to hold.
+Indeed, all that is needed is the intrinsic dimension $m$--- in order to properly tune $r$ and $\rho$--- and otherwise $\wh{f}$ and $\wh{\varphi}$ are computed without regard to $\Xset$.
+In contrast, the penalty in~\eqref{eqn:thin_plate_spline} would have to be specially tailored to work in this setting, revealing another advantage of Laplacian smoothing compared to thin-plate splines.
 
@@
 
+\textcolor{red}{(TODO)}
+
 \section{DISCUSSION}
 \label{sec:discussion}
+
+In this work, we have shown that Laplacian smoothing over a neighborhood graph is optimal, for various statistical problems and under various assumptions.
+
+There are still several extensions we feel are worth pursuing.
+In practice, it is more common to use $k$NN graphs than neighborhood graphs, due to the guaranteed connectivity and sparsity of the former; we strongly suspect that by building on the work of~\cite{calder2019}, one can show that our main results all hold when the $k$NN graph is used.
+In another direction, one can also generalize Laplacian smoothing by replacing the penalty $f^T \Lap_{n,r} f$ with $f^T \Lap_{n,r}^k f$ for some $k > 1$.
+The hope is that this adapted procedure would be minimax optimal over the previously mentioned higher-order Sobolev classes $H^k(\Xset)$.
+In the very special case of $k = 2$ and $d \leq 1 2$, we can show that this is indeed the case, but the general story--for all combinations of $k$ and $d$---remains beyond our reach.
 

### 2020-10-14 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing4.tex
Begun a new draft---xxx_laplacian_smoothing4--- in response to feedback from Ryan and Siva.

--- 2020-10-09 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing3.tex
@@
 	In this paper we study the statistical properties of Laplacian smoothing, a graph-based approach to non-parametric regression.
-We show that under standard regularity conditions, the Laplacian smoothing estimator $\wh{f}$, and a test using the statistic $\wh{T} = \frac{1}{n}\|\wh{f}\|_2^2$, attain the minimax optimal estimation and testing rates of convergence over the first-order Sobolev class $H^1(\Xset)$, where $\Xset \subset \Rd$ and $1 \leq d \leq 4$.
+Under standard regularity conditions, we establish upper bounds on the error of the Laplacian smoothing estimator $\wh{f}$, and a goodness-of-fit test based on $\wh{f}$.
+These upper bounds match the minimax optimal estimation and testing rates of convergence over the first-order Sobolev class $H^1(\Xset)$, where $\Xset \subset \Rd$ and $1 \leq d < 4$; in the estimation problem, they are within a $\log n$ factor of optimal when $d = 4$.
 Additionally, we prove that Laplacian smoothing is manifold adaptive: if $\Xset$ is an $m$ dimensional manifold embedded in $\Reals^d$ with $m \ll d$, then the error rate of Laplacian smoothing methods depends only on the intrinsic dimension $m$ and not on the ambient dimension $d$.
-We support our theory with simulations.
+Empirically, we show that Laplacian smoothing appears not to be minimax optimal when $d > 4$.
 \end{abstract}
@@
 There are known extremely fast and reliable solvers for exactly this problem (see e.g.
-the seminal papers by \cite{spielman2011,spielman2013,spielman2014}, or the review article~\cite{vishnoi2012} and references therein).
+the seminal papers of \cite{spielman2011,spielman2013,spielman2014}, or the monograph by~\cite{vishnoi2012} and references therein).
 	\item \emph{Generality.} Laplacian smoothing is well defined whenever one can associate a graph with some observed responses.
 This generality lends itself to many different data modalities: for instance text and image classification, to name only two \citep{kondor2002, belkin03a,belkin2006} \textcolor{red}{(TODO)}: Add a bunch more references here.
-	\item \emph{Supervision.} Although we study Laplacian smoothing in the supervised learning setup~\eqref{eqn:signal_plus_noise_model}, it can be easily adapted to the semi-supervised or unsupervised settings \citep{nadler09,slepcev17,dunlop2020,calder2019b}.
+	\item \emph{Weak supervision.} Although we study Laplacian smoothing in the supervised learning setup~\eqref{eqn:signal_plus_noise_model}, it can be easily adapted to the semi-supervised or unsupervised settings \citep{nadler09,slepcev17,dunlop2020,calder2019b}.
 \end{itemize}
@@
 
-\textcolor{red}{(TODO)}: Make Table 1.
+\begin{table}
+\begin{center}
+\begin{tabular}{p{.1\textwidth} | p{.14\textwidth} p{.12\textwidth} }
+	Dimension & Laplacian smoothing \eqref{eqn:laplacian_smoothing} & Thin-plate splines \eqref{eqn:thin_plate_spline} \\
+	\hline
+	$d = 1$ & $n^{-2/3}$ & \textcolor{red}{$n^{-2/3}$} \\
+	$d = 2,3$ & $n^{-2/(2 + d)}$ & \textcolor{red}{$1$} \\
+	$d = 4$ & $n^{-2/6}\log n$ & \textcolor{red}{$1$} \\
+	$d \geq 5$ & \textcolor{red}{(?)} & \textcolor{red}{$1$} \\
+\end{tabular}
+\end{center}
+\caption{Summary of estimation rates over first-order Sobolev balls; rates in black are new results of this paper.
+When $\Xset$ is an $m$ dimensional manifold embedded in $\Rd$, all rates hold with $d$ replaced by $m$.}
+\label{tbl:estimation_rates}
+\end{table}
+
+\begin{table}
+\begin{center}
+	\begin{tabular}{p{.1\textwidth} | p{.14\textwidth} p{.12\textwidth} }
+		Dimension & Laplacian smoothing \eqref{eqn:laplacian_smoothing_test} & Thin-plate splines \eqref{eqn:thin_plate_spline} \\
+		\hline
+		$d = 1$ & $n^{-4/5}$ & \textcolor{red}{$n^{-4/5}$} \\
+		$d = 2,3$ & $n^{-4/(4 + d)}$ & \textcolor{red}{$n^{-1/2}$} \\
+		$d \geq 4$\footnotemark & $n^{-1/4}$ & \textcolor{red}{$n^{-1/4}$}
+	\end{tabular}
+\end{center}
+\caption{Summary of testing rates over first-order Sobolev balls; rates in black are new results of this paper.
+When $\Xset$ is an $m$ dimensional manifold embedded in $\Rd$, all rates hold with $d$ replaced by $m$.
+Rates for $d \geq 4$ assume $f_0 \in \Leb^4(\Xset,M)$.}
+\label{tbl:testing_rates}
+\end{table}
+\footnotetext{Assuming $f_0 \in \Leb^4(\Xset,M)$.}
 
@@
 \end{equation*}
-where $\norm{\cdot} = \norm{\cdot}_{\Rd}$ is the Euclidean norm.
+where $\norm{\cdot}$ is the Euclidean norm in $\Rd$.
 Then the degree matrix $D$ is the $n \times n$ diagonal matrix with entries $D_{ii} = \sum_{j = 1}^{n}W_{ij}$, and the graph Laplacian can be written as
@@
 \paragraph{Sobolev spaces.}
-Now that we've defined the neighborhood graph Laplacian $\Lap_{n,r}$, we step away from graph-based methods for a moment, to briefly recall some very classical results regarding minimax rates over Sobolev classes.
+Now that we've defined the neighborhood graph Laplacian $\Lap_{n,r}$, we step away from graph-based methods for a moment, to briefly recall some classical results regarding minimax rates over Sobolev spaces.
 
@@
 \end{equation*}
+For further details regarding Sobolev spaces see e.g \citep{evans10, leoni2017}.
 
 \paragraph{Minimax rates.}
-To carry out a minimax analysis of Sobolev spaces, one must impose some regularity conditions on the design distribution $P$. 
+To carry out a minimax analysis of regression in Sobolev spaces, one must impose some regularity conditions on the design distribution $P$. 
 \begin{enumerate}[label=(P\arabic*)]
@@
 
-Under~\ref{asmp:bounded_lipschitz_density} and~\ref{asmp:domain}, the minimax estimation rate over Sobolev balls is (see e.g. \citep{tsybakov2008_book})
+Under~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density}, the minimax estimation rate over Sobolev balls is (see e.g. \citep{tsybakov2008_book}),
 \begin{equation}
 \label{eqn:sobolev_space_estimation_minimax_rate}
-\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{d/(2 + d)}n^{-2/(2 + d)}
+\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{d/(2 + d)}n^{-2/(2 + d)}.
 \end{equation}
@@
 As nonparametric hypothesis testing is (comparatively) less familiar than nonparametric estimation, we briefly summarize the main idea before stating the optimal error rate.
-In the goodness-of-fit testing problem, we ask for a test function which can distinguish between the null and alternative hypotheses
+In the goodness-of-fit testing problem, we ask for a test function---formally, a Borel measurable function $\phi$ that takes values in $\{0,1\}$--- which can distinguish between the null and alternative hypotheses
 \begin{equation}
-\mathbf{H}_0: f_0 = 0, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{0\}
+\mathbf{H}_0: f_0 = f_0^{\star}, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{f_0^{\star}\}.
 \end{equation} 
-where $\mc{F}$ is some class of functions.
+Typically, the null hypothesis $f_0 = f_0^{\star} \in \mc{F}$ reflects the absence of interesting structure, and $\mc{F} \setminus  \{f_0^{\star}\}$ is a set of smooth departures from this null.
+To fix ideas, as in \cite{ingster2009} we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\star} = 0$ and $\mc{F} = H^1(\Xset,M)$ is a first-order Sobolev ball; this is without loss of generality.
+
 The minimax critical radius is the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{F}_{\epsilon} := \mc{F} \cap \{f: \|f\|_{\Leb^2(\Xset)} \geq \epsilon\}$:
 \begin{equation*}
-\epsilon(\mc{F}) := \inf\Biggl\{\epsilon > 0: \inf_{\phi} \biggl[\Ebb_0[\phi] +  \sup_{f_0 \in \mc{F}} \Ebb_{f_0}[1 - \phi]\biggr] \leq 2\alpha\Biggr\}
+\epsilon(\mc{F}) := \inf\Biggl\{\epsilon > 0: \inf_{\phi} \biggl[ \sup_{f_0 \in \mc{F}_{\epsilon}} \Ebb_{f_0}[1 - \phi]\biggr] \leq \alpha\Biggr\}
 \end{equation*} 
-where $\Ebb_{f_0}[\cdot]$ is the expectation under the regression function $f_0$. See \cite{ariascastro2018} for a more extended discussion of the minimax paradigm in nonparametric testing.
+where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation under the regression function $f_0$.\footnote{\textcolor{red}{(For Siva)}: Clearly, $\epsilon$ depends on $\alpha$.
+However we adopt the typical convention of treating $\alpha$ as a small but fixed positive constant; hence it will not affect the testing error rates, and we suppress it notationally.} See \citep{ingster82, ingster87, ingster2012, ariascastro2018} for a more extended treatment of the minimax paradigm in nonparametric testing.
 
@@
 \end{equation}
-When $d \geq 4$ the functions in $H^1(\Xset)$ are very irregular---formally speaking $H^1(\Xset)$ does not continuously embed into $\Leb^4(\Xset)$ when $d \geq 4$---and the minimax rates in this regime are unknown.
+When $d \geq 4$ the functions in $H^1(\Xset)$ are very irregular---formally speaking $H^1(\Xset)$ does not continuously embed into $\Leb^4(\Xset)$ when $d \geq 4$---and the minimax testing rates in this regime are unknown.
 
@@
 \paragraph{Estimation error of Laplacian smoothing.} 
-Under these conditions, the Laplacian smoothing estimator $\wh{f}$ achieves, up to constants, the minimax risk $n^{-2/(2 + d)}$ over all $f_0 \in H^1(\Xset,M)$.
+Under these conditions, the Laplacian smoothing estimator $\wh{f}$ achieves, up to constants, the minimax risk $n^{-2/(2 + d)}$ over $H^1(\Xset,M)$.
 This statement will hold whenever the graph $G_{n,r}$ is computed with radius $r$ in the following range.
@@
 	\begin{equation*}
-	\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq r \leq n^{-3/(4 + 2d)} M^{(d - 4)/(4 + 2d)} \wedge c_0.
+	C_0\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq r \leq n^{-3/(4 + 2d)} M^{(d - 4)/(4 + 2d)} \wedge c_0.
 	\end{equation*}
@@
 	\label{thm:laplacian_smoothing_estimation1}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d < 4$ and $f_0 \in H^1(\Xset,M)$.
+	Let $f_0 \in H^1(\Xset,M)$ with $d < 4$.
 Suppose that that the neighborhood graph $G_{n,r}$ is computed with a radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_estimation},  and the Laplacian smoothing estimator $\wh{f}$ with $\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$.
-It holds that
+With probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c M^{d/(2d + 4)} n^{d/(2+d)})$, it holds that
 	\begin{equation*}
-	\Bigl\|\wh{f} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{d/(2 + d)} n^{-2/(2 + d)}
+	\Bigl\|\wh{f} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(2 + d)} n^{-2/(2 + d)}.
 	\end{equation*}
-	with probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c M^{d/(2d + 4)} n^{d/(2+d)})$.
 \end{theorem}
@@
 \cite{sadhanala16} argue that their analysis is tight, and that it is the estimator, rather than the analysis, which is deficient when $d \geq 5$.
-Formally proving such a claim turns out to be harder in the random design setting than in the fixed design setting; although we suspect a similar claim is true, we cannot prove it, and instead investigate the matter empirically in Section~\ref{sec:simulations}.
+Formally proving such a claim turns out to be harder in the random design setting than in the fixed design setting; however we conjecture a similar claim is true, and investigate the matter empirically in Section~\ref{sec:simulations}.
 	\item The lower bound on $r$ imposed by~\ref{asmp:ls_kernel_radius_estimation} is  compatible with practice, where by far the most common choice of radius is the connectivity threshold $r \asymp (\log(n)/n)^{1/d}$, chosen to make the graph $G_{n,r}$ as sparse as possible while still being connected, for maximum computational efficiency.
@@
 In fact, the threshold $\wh{t}_b$ is precisely the right choice to control the Type I error of $\wh{\varphi}$; as we show in Lemma~\ref{lem:ls_fixed_graph_testing} in the supplementary material,
-\begin{equation*}
+\begin{equation}
+\label{eqn:type_I_error}
 \Ebb_0\bigl[\wh{\varphi}\bigr] \leq \frac{1}{b^2}.
-\end{equation*}
+\end{equation}
 Theorem~\ref{thm:laplacian_smoothing_testing} establishes that this test also has small Type II error, uniformly over all $f_0$ separated from $0$ by at least the critical radius $\epsilon\bigl(H^1(\Xset,M)\bigr)$ given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
@@
 	\begin{equation*}
-	\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq r \leq M^{(16 - d)/(8 + 2d)}n^{(d - 20)/(32 + 8d)} \wedge c_0.
+	C_0\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq r \leq M^{\frac{(16 - d)}{8 + 2d}}n^{\frac{d - 20}{32 + 8d}} \wedge c_0.
 	\end{equation*}
 \end{enumerate}
-
+Note that when $d < 4$, as long as $M$ is not too small there always exists a choice of $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}.
 \begin{theorem}
 	\label{thm:laplacian_smoothing_testing}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d < 4$ and $f_0 \in H^1(\Xset,M)$ for $M \leq n^{(4 - d)/(4 + d)}$.
+	Let $f_0 \in H^1(\Xset,M)$ for $M \leq n^{(4 - d)/(4 + d)}$ and $d < 4$.
 Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}, and the Laplacian smoothing test $\wh{\varphi}$ with $\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$, and $b \geq 1$.
@@
 setting $\rho = 0$, the test $\wh{\varphi}$ achieves (up to constants) this lower bound.
-That is, for any $f_0 \in \Leb^4(\Xset,M)$ such that $\|f_0\|_{\Leb^2(\Xset)}^2 \geq C b^2n^{-1/2}$, we have
+That is, for any $f_0 \in \Leb^4(\Xset,M)$ such that $\|f_0\|_{\Leb^2(\Xset)}^2 \geq C b^2n^{-1/2}$, we have both
 	\begin{equation}
@@
 	\end{equation} 
+	and $\Ebb_0[\wh{\varphi}] \leq 1/b^2$.
 	\item To compute the data-dependent threshold $\wh{t}_b$, one must know each of the eigenvalues $\lambda_1,\ldots,\lambda_n$.
 Computing all $n$ of these eigenvalues is far more expensive---on the order of $n^3$ operations---than computing $\wh{T}$.
-That being said, in practice we would not recommend using $\wh{t}_b$ anyway: the threshold is justified by asymptotic theory, but in practice $n$ is always finite, and using such thresholds can result in substantial loss of efficiency.
-We would instead make the standard recommendation to calibrate via permutation.
+That being said, in practice we would not recommend using $\wh{t}_b$ anyway, as the threshold is derived through a (potentially loose) application of Chebyshev's inequality which can lead to a loss in effiency.
+Instead we make the standard recommendation to calibrate via permutation \citep{hoeffding1952}.
 \end{itemize}
@@
 The key point is that the discretization imposed by the graph $G_{n,r}$ turns out to be a blessing.
-The problem with~\ref{eqn:thin_plate_spline} is that the function class $H^1(\Xset)$ is far too large when $d \geq 2$; we know from the Sobolev embedding theorem \cite{evans10} that it does not continuously embed into any Holder space, and that it is not an Reproducing Kernel Hilbert Space.
+The problem with~\eqref{eqn:thin_plate_spline} is that the function class $H^1(\Xset)$ is far too large when $d \geq 2$; we know from the Sobolev embedding theorem \cite{evans10} that it does not continuously embed into any H\"{o}lder space, and that it is not an Reproducing Kernel Hilbert Space.
 On the other hand~\eqref{eqn:laplacian_smoothing} is a finite dimensional (if growing linearly with $n$) problem---as a result $\wh{f}$ has far less capacity to overfit than does $\wt{f}$.
@@
 where $|f|_{H^k(\Xset)}^2$ is the order-$k$ Sobolev seminorm, which measures the size of the order-$k$ partial derivatives in $L^2(\Xset)$ norm.
-Assuming $f_0 \in H^k(\Xset)$ for some $2k > d$, the degree-$k$ thin plate spline has error on the order of $n^{-2k/(2k + d)}$ \cite{vandergeer2000}, which is the minimax optimal rate.
+Assuming $f_0 \in H^k(\Xset)$ for some $2k > d$, the degree-$k$ thin plate spline has error on the order of $n^{-2k/(2k + d)}$ \citep{vandergeer2000}, which is the minimax optimal rate.
 Of course, assuming $f_0 \in H^k(\Xset)$ for some $2k > d$ is a rather strong assumption which may not be warranted, but at present we do not know whether (adaptations of) Laplacian smoothing on neighborhood graphs can achieve these faster rates.
@@
 The previous discussion shows some surprising strengths of the Laplacian smoothing estimator, but it also highlights the difficulty of our analysis task.
-A natural way to analyze $\wh{f}$ would be to use~\eqref{eqn:seminorm_consistency} to establish a coupling with $\wt{f}$, but this will not do the job: first, proving such a coupling is tricky---note that~\eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:thin_plate_spline} is $H^1(\Xset)$---and second, as we've just established we actually would like to bound $\wh{f}$ in settings where we know $\wt{f}$ fails.
+A natural way to analyze $\wh{f}$ would be to use~\eqref{eqn:seminorm_consistency} to establish a coupling with $\wt{f}$, but this will not do the job: first, establishing such a coupling is tricky---note that~\eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:thin_plate_spline} is $H^1(\Xset)$---and second, as we've just established we actually would like to bound $\wh{f}$ in settings where we know $\wt{f}$ fails.
 
@@
 \begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{2}
 	\item
@@
 	\label{lem:graph_sobolev_seminorm}
-	For any $f \in H^1(\Xset)$, 
+	For any $f \in H^1(\Xset)$, with probability at least $1 - \delta$, it holds that
 	\begin{equation}
 	\label{eqn:graph_sobolev_seminorm}
-	f^T \Lap_{n,r} f \leq \frac{(1 + L_p)^2 \sigma_K}{\delta} n^2 r^{d + 2} |f|_{H^1(X)}^2
+	f^T \Lap_{n,r} f \leq \frac{(1 + L_p)^2 \sigma_K}{\delta} n^2 r^{d + 2} |f|_{H^1(X)}^2.
 	\end{equation}
-	with probability at least $1 - \delta$.
 \end{lemma}
@@
 	\label{lem:neighborhood_eigenvalue}
-	For any neighborhood graph radius $r$ which satisfies~\ref{asmp:radius1}, it holds that
+	For any neighborhood graph radius $r$ which satisfies~\ref{asmp:radius1}, with probability at least $1 - C_1n\exp(-c_1nr^d)$ it holds that
 	\begin{equation}
@@
 	\end{equation}
-	with probability at least $1 - C_1n\exp(-c_1nr^d)$.
 \end{lemma}
@@
 The parameter $\rho$ is then tuned to minimize the sum of bias and variance, in the usual manner.
-
-\textcolor{red}{(For Siva)}: I think the below paragraph as currently written is not good enough.
-My question(s) for you: do you agree, and if so, do you think it can be fixed by rewriting or should be shelved entirely?
 
@@
 The minimax rates $n^{-2/(2 + d)}$ (in estimation) and $n^{-4/(4 + d)}$ (in testing) suffer from the curse of dimensionality.
-One way to overcome this is the \emph{manifold hypothesis}: roughly speaking, that the data $X_1,\ldots,X_n$ lie on a submanifold $\Xset$ of $\Rd$ that has intrinsic dimension $1 \leq m \leq d$.
+However, in practice it is often reasonable to assume a \emph{manifold hypothesis}: roughly speaking, that the data $X_1,\ldots,X_n$ lie on a submanifold $\Xset$ of $\Rd$ that has intrinsic dimension $1 \leq m \leq d$.
 Under such an assumption, it is known \citep{bickel2007, ariascastro2018} that the optimal rates over the Lipschitz classes $C^1(\Xset)$ now scale like $n^{-2/(2 + m)}$ (for estimation) and $n^{-4/(4 + m)}$ (for testing), which can be much smaller than the full-dimensional error rates when $m \ll d$.
@@
 	\begin{equation*}
-	\biggl(\frac{\log n}{n}\biggr)^{1/m} \leq r \leq n^{-3/(4 + 2m)} M^{(m - 4)/(4 + 2m)} \wedge c_3.
+	C_3\biggl(\frac{\log n}{n}\biggr)^{1/m} \leq r \leq n^{-3/(4 + 2m)} M^{(m - 4)/(4 + 2m)} \wedge c_3.
 	\end{equation*}
@@
 	\label{thm:laplacian_smoothing_estimation_manifold}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $m < 4$ and $f_0 \in H^1(\Xset,M)$.
+	Let $f_0 \in H^1(\Xset,M)$ with $m < 4$.
 Suppose that that the neighborhood graph $G_{n,r}$ is computed with a radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_estimation_manifold},  and the Laplacian smoothing estimator $\wh{f}$ with $\rho = M^{-4/(2 + m)} (nr^{m + 2})^{-1} n^{-2/(2 + m)}$.
-It holds that
+With probability at least $1 - \delta -  C_4n\exp(-c_4nr^m) - \exp(-c M^{m/(2m + 4)} n^{m/(2+m)})$, it holds that
 	\begin{equation*}
-	\Bigl\|\wh{f} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{m/(2 + m)} n^{-2/(2 + m)}
+	\Bigl\|\wh{f} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2m/(2 + m)} n^{-2/(2 + m)}.
 	\end{equation*}
-	with probability at least $1 - \delta -  C_4n\exp(-c_4nr^m) - \exp(-c M^{m/(2m + 4)} n^{m/(2+m)})$.
 \end{theorem}
 
+We require the following range of $r$ for testing.
 \begin{enumerate}[label=(R\arabic*)]
@@
 	\begin{equation*}
-	\biggl(\frac{\log n}{n}\biggr)^{1/m} \leq r \leq M^{\frac{(16 - m)}{8 + 2m}}n^{\frac{m - 20}{32 + 8m}} \wedge c_3.
+	C_3\biggl(\frac{\log n}{n}\biggr)^{1/m} \leq r \leq M^{\frac{(16 - m)}{8 + 2m}}n^{\frac{m - 20}{32 + 8m}} \wedge c_3.
 	\end{equation*}
 \end{enumerate}
+Similarly to the full-dimensional setting, when $m < 4$ there exists a choice of $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing_manifold} so long as $M$ is not too small.
 
@@
 	\label{thm:laplacian_smoothing_testing_manifold}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $m < 4$ and $f_0 \in H^1(\Xset,M)$ for $M \leq n^{(4 - m)/(4 + m)}$.
+	Let $f_0 \in H^1(\Xset,M)$ for $M \leq n^{(4 - m)/(4 + m)}$ and $m < 4$.
 Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing_manifold}, and the Laplacian smoothing test $\wh{\varphi}$ with $\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$, and $b \geq 1$.
@@
 
-\textcolor{red}{(TODO)}
+In Figure~\ref{fig:fig1}, we empirically investigate the mean squared error of the Laplacian smoothing estimator, for different values of dimension $d$.
+In these experiments, $X_1,\ldots,X_n$ are sampled uniformly over $\Xset = [-1,1]^d$, and the regression function $f_0(x) = \cos(m \pi x)$, where $m = 6$ for $d = 1$, $m = 2$ for $d = 2$, and $m = 1$ for $d \geq 3$.
+For reference, we compare Laplacian smoothing to a $k$NN estimator, which is known to achieve the minimax optimal rate $n^{-2/(2+d)}$ for all $d \geq 1$.
+
+When $d = 4$ this continues to be the case for $k$NN regression, but not for Laplacian smoothing.
+In this case, the \textcolor{red}{expected slope} for Laplacian smoothing over this range of $n$---based on the upper bound $\log(n)n^{-1/4}$ derived in Theorem~\ref{thm:laplacian_smoothing_estimation1}---is $.18$.
+
+\begin{figure}[thb]
+	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_2d.pdf}
+	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_3d.pdf}
+	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_4d.pdf}
+	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_5d.pdf}
+
+	\caption{Comparison of Laplacian smoothing and $k$NN regression, as a function of sample size $n$.
+Plots are on the log-log scale, MSEs are averaged over 5 iterations, with both methods tuned for optimal MSE.
+When $d = 1,2,3$, empirically both methods are achieving or exceeding the minimax rate.}
+	\label{fig:fig1}
+\end{figure}
 

### 2020-10-16 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing.tex
Final draft for AISTATS submission.

@@
-\documentclass{article}
-\usepackage{amsmath}
-\usepackage{amsfonts, amsthm, amssymb}
+\documentclass[twoside]{article}
+\usepackage{aistats2021}
+\usepackage{amsmath, amsfonts, amsthm, amssymb}
 \usepackage{graphicx}
@@
 \usepackage[shortlabels]{enumitem}
-\usepackage{fullpage}
 \usepackage{mathtools}
 \usepackage{tikz}
+\usepackage{verbatim}
 
@@
 \DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
+\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}
+
+% Make a widecheck symbol (thanks, Stack Exchange!)
+\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
+\DeclareFontShape{U}{mathx}{m}{n}{
+	<5> <6> <7> <8> <9> <10>
+	<10.95> <12> <14.4> <17.28> <20.74> <24.88>
+	mathx10
+}{}
+\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
+\DeclareFontSubstitution{U}{mathx}{m}{n}
+\DeclareMathAccent{\widecheck}{0}{mathx}{"71}
+% widecheck made
 
@@
 \begin{document}
-\title{Minimax optimal Laplacian smoothing}
-\author{Alden Green}
-\date{\today}
-\maketitle
 
-\textbf{Disclaimer:} This largely follows the same setup as in the draft I have written which combines Laplacian smoothing and Laplacian eigenmaps results.
-That draft --- document name \textit{graph\_regression2} --- should for the moment be considered obsolete.
+\twocolumn[
 
-\section{Introduction}
+	\aistatstitle{Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs}
 
-\textbf{(1)} Laplacian smoothing is a graph-based approach to nonparametric regression.
+	\aistatsauthor{Alden Green, Sivaraman Balakrishnan, and Ryan J. Tibshirani}
 
-\begin{itemize}
-	\item In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Rd$, and
+	\aistatsaddress{Carnegie Mellon University}]
+
+\begin{abstract}
+	In this paper we study the statistical properties of Laplacian smoothing, a graph-based approach to non-parametric regression.
+Under standard regularity conditions, we establish upper bounds on the error of the Laplacian smoothing estimator $\wh{f}$, and a goodness-of-fit test based on $\wh{f}$.
+These upper bounds match the minimax optimal estimation and testing rates of convergence over the first-order Sobolev class $H^1(\Xset)$, where $\Xset \subset \Rd$ and $1 \leq d < 4$; in the estimation problem, they are within a $\log n$ factor of optimal when $d = 4$.
+Additionally, we prove that Laplacian smoothing is manifold adaptive: if $\Xset$ is an $m$ dimensional manifold embedded in $\Reals^d$ with $m \ll d$, then the error rate of Laplacian smoothing methods depends only on the intrinsic dimension $m$ and not on the ambient dimension $d$.
+\end{abstract}
+
+\section{INTRODUCTION}
+In the nonparametric regression problem, we observe samples $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to the signal plus noise model
 \begin{equation}
-	\label{eqn:random_design_regression}
-	Y_i = f_0(X_i) + \varepsilon_i
+\label{eqn:signal_plus_noise_model}
+Y_i = f_{0}(X_i) + \varepsilon_i,
 \end{equation}
 with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise.
-Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present.
-	\item In graph-based nonparametric regression, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$.
-The neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$.
-The $n \times n$ weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $\mathbf{W}_{ij}$ are given by
+Our goal is to perform statistical inference on the unknown regression function $f_0$, by which we mean either \emph{estimating} $f_0$ or more simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present.
+
+The Laplacian smoothing estimator $\wh{f}$ \citep{smola2003} is a penalized least squares estimator, defined on a graph.
+Letting $G = \bigl(\{1,\ldots,n\},W\bigr)$ be a weighted graph over vertices $\{1,\ldots,n\}$ which we associate with the sample points $\{X_1,\ldots,X_n\}$, the estimator $\wh{f}$ is given by
+\begin{equation}
+\label{eqn:laplacian_smoothing}
+\wh{f} :=  \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap f \biggr\}.
+\end{equation}
+Here $\Lap$ is the graph Laplacian matrix (defined in Section~\ref{sec:problem_setup_and_background}), $G$ is typically a geometric graph (such as a $k$NN or neighborhood graph) and the penalty
+\begin{equation*}
+f^T \Lap f = \frac{1}{2} \sum_{i,j = 1}^{n} W_{ij}\bigl(f_i - f_j\bigr)^2
+\end{equation*}
+thus encourages $\wh{f}_i \approx \wh{f}_j$ when $X_i \approx X_j$. Assuming~\eqref{eqn:laplacian_smoothing} is a reasonable estimator of $f_0$, the statistic
+\begin{equation}
+\label{eqn:laplacian_smoothing_test}
+\wh{T} := \frac{1}{n}\bigl\|\wh{f}\bigr\|_2^2 
+\end{equation}
+is in turn a reasonable way to decide if $f_0 = 0$.
+
+Of course there exist many methods for nonparametric regression \citep{gyorfi2006,wasserman2006,tsybakov2008_book}, but Laplacian smoothing has its own set of advantages. For instance:
+\begin{itemize}
+	\item \emph{Computational ease.} Laplacian smoothing is fast, easy, and stable to compute.
+The estimate $\wh{f}$ can be computed by solving a symmetric, diagonally dominant linear system.
+There are known extremely fast and reliable solvers for exactly this problem (see e.g.
+the seminal papers of \cite{spielman2011,spielman2013,spielman2014}, or the monograph by~\cite{vishnoi2012} and references therein).
+	\item \emph{Generality.} Laplacian smoothing is well defined whenever one can associate a graph with some observed responses.
+This generality lends itself to many different data modalities: for instance text and image classification, to name only two \citep{kondor2002, belkin03a,belkin2006}.
+	\item \emph{Weak supervision.} Although we study Laplacian smoothing in the supervised learning setup~\eqref{eqn:signal_plus_noise_model}, it can be easily adapted to the semi-supervised or unsupervised settings \citep{nadler09,slepcev17,dunlop2020,calder2019b}.
+\end{itemize}
+
+For these reasons, a body of work has emerged analyzing the statistical properties of Laplacian smoothing, and graph-based methods more generally.
+Roughly speaking, these works can be divided into two categories, based on the perspective they adopt.
+\begin{itemize}
+	\item In the \emph{fixed design perspective}, the design points $\{X_1,\ldots,X_n\}$ and the graph $G$ are treated as fixed, and inference is carried out on the evaluations $\bigl(f_0(X_1),\ldots,f_0(X_n)\bigr) \in \Reals^n$.\footnote{To be clear, while in this setting $G$ may be a geometric graph, in principle it could be any graph over $\{1,\ldots,n\}$.} In this context, tight upper bounds have been derived on the error of graph-based estimators \citep{wang2016, sadhanala16,sadhanala17,kirichenko2017,kirichenko2018} and tests \citep{sharpnack2013,sharpnack2013b,sharpnack2015}, which certify that such procedures are \emph{optimal} over ``function'' classes $\mc{F} \subseteq \Reals^n$.
+The downside of this work is that it may be unnatural to assume a priori that the graph $G$ is fixed, or that the evaluations $\bigl(f_0(X_1),\ldots,f_0(X_n)\bigr)$ belong to some discrete ``function'' class.
+	\item In the arguably more natural \emph{random design perspective},  one assumes the design points $X_1,\ldots,X_n$ are sampled independently from some distribution $P$ supported on a domain $\Xset \subset \Rd$.
+Inference is drawn on the regression function $f_0: \Xset \to \Reals$, which is typically assumed to be smooth in some continuum sense, e.g.
+it possesses a first derivative bounded in $L^{\infty}$ (H\"{o}lder) or $\Leb^2$ (Sobolev) norm.
+To conduct graph-based inference, the user first builds a neighborhood graph over the random design points---so that $W_{ij}$ is large iff $X_i$ and $X_j$ are close in say Euclidean distance---and then solves e.g.~\eqref{eqn:laplacian_smoothing} or~\eqref{eqn:laplacian_smoothing_test}.
+In this context, various graph-based procedures have been shown to be \emph{consistent}: as $n \to \infty$, they converge to a \emph{continuum} limit (see \citep{belkin07,trillos2018,vonluxburg2008} among others).
+However, until recently such statements were not accompanied by error rates, and such error rates as have been proved \citep{lee2016,trillos2020} are not optimal over continuum function spaces, such as H\"{o}lder or Sobolev classes.
+\end{itemize}
+Hereafter, we will adopt the random design perspective, and seek to answer the following question:
+\begin{quote}{When $f_0$ is assumed to smooth in a continuum sense, does Laplacian smoothing achieve optimal performance for estimation and goodness-of-fit testing?} 
+\end{quote}
+On the one hand, this is no small question---rather, it is arguably the fundamental question of nonparametric regression, and without an answer one cannot fully compare the statistical properties of Laplacian smoothing to its competitors.
+On the other hand, it seems difficult to answer: there is a fundamental gap between the \emph{discrete} smoothness imposed by the penalty $f^T \Lap f$ in~\eqref{eqn:laplacian_smoothing} and the \emph{continuum} smoothness assumed on $f_0$, and to obtain sharp upper bounds we will need to bridge this gap in a tight sense.
+
+\section{SUMMARY OF RESULTS}
+
+\paragraph{Advantages of the discrete approach.} In light of this last point, it is worth asking whether there is any \emph{statistical} advantage to solving a discrete problem such as~\eqref{eqn:laplacian_smoothing} (setting aside computational considerations for the moment).
+After all, we could have instead solved the following variational problem:
+\begin{equation}
+\label{eqn:thin_plate_spline}
+\wt{f} := \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \cdot \int_{\Xset} \|\nabla f(x)\|^2 \,dx \biggr\},
+\end{equation}
+with the minimum computed over all functions $f$ for which the criterion is finite--or used the statistic
+\begin{equation}
+\label{eqn:thin_plate_spline_test}
+\wt{T} := \|\wt{f}\|_n^2.
+\end{equation}
+The penalty in~\eqref{eqn:thin_plate_spline} leverages the assumption that $f_0$ possesses a bounded derivative in a seemingly natural manner.
+Indeed, for this reason the estimator $\wt{f}$ (and to a lesser extent, the statistic $\wt{T}$) are well known--- in the univariate case, $\wt{f}$ is simply a \emph{smoothing spline}, and when $d > 1$ it is an example of a \emph{thin-plate spline}---and their statistical properties are well-understood \citep{vandergeer2000, liu2019}.
+Moreover, there exist strong connections between smoothing/thin-plate splines and Laplacian smoothing: as we show later on in~\eqref{eqn:seminorm_consistency}, in a certain sense the Laplacian smoothing problem~\eqref{eqn:laplacian_smoothing} can be seen as a discrete and noisy approximation to~\eqref{eqn:thin_plate_spline}.
+At first blush, this suggests that Laplacian smoothing should at best inherit the statistical properties of smoothing/thin-plate splines, and at worst may have meaningfully larger error.
+
+However, as we shall see the actual story is quite different: remarkably, Laplacian smoothing enjoys optimality properties in cases where the thin-plate spline problem~\eqref{eqn:thin_plate_spline} is not even well-posed.
+Tables~\ref{tbl:estimation_rates} and~\ref{tbl:testing_rates} summarize the gap.
+As we establish in Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing_manifold}, when computed over an appropriately formed neighborhood graph, Laplacian smoothing estimators and tests are minimax optimal over first-order \emph{continuum} Sobolev balls.
+This holds true either when $\Xset \subset \Rd$ is a full-dimensional domain and $d = 1,2$ or $3$, or when $\Xset$ is a submanifold of $\Rd$ of intrinsic dimension $m = 1,2,$ or $3$.
+Additionally, the estimator $\wh{f}$ is nearly minimax optimal (to within a $(\log n)^{1/3}$ factor) when $d$ (or $m$) = $4$.
+By contrast, smoothing splines are optimal only when $d$ (or $m$) = $1$.
+In higher dimensions, technically speaking the problem~\eqref{eqn:thin_plate_spline} is not even well-posed, and any practical implementation of the estimator $\wt{f}$ will be inconsistent \citep{green93}.
+Hence, our results favorably distinguish Laplacian smoothing methods from their natural variational alternatives.
+\begin{table}
+	\begin{center}
+		\begin{tabular}{p{.1\textwidth} | p{.14\textwidth} p{.12\textwidth} }
+			Dimension & Laplacian smoothing \eqref{eqn:laplacian_smoothing} & Thin-plate splines \eqref{eqn:thin_plate_spline} \\
+			\hline
+			$d = 1$ & ${\bf n^{-2/3}}$ & \textcolor{red}{${\bf n^{-2/3}}$} \\
+			$d = 2,3$ & ${\bf n^{-2/(2 + d)}}$ & \textcolor{red}{$1$} \\
+			$d = 4$ & ${\bf n^{-1/3}} (\log n)^{1/3}$ & \textcolor{red}{$1$} \\
+			$d \geq 5$  & $(\log n/n)^{4/(3d)}$ &\textcolor{red}{$1$} \\
+		\end{tabular}
+	\end{center}
+	\caption{Summary of estimation rates over first-order Sobolev balls.
+Rates in black are new results of this paper, and bolded rates are minimax optimal.
+Although we suppress it for conciseness, in all cases the dependence of the error rate on the radius of the Sobolev ball is also optimal.
+For thin-plate splines, ``\textcolor{red}{$1$}'' indicates that the estimator is inconsistent.
+When $\Xset$ is an $m$ dimensional manifold embedded in $\Rd$, all rates hold with $d$ replaced by $m$, without any change in the method in the case of Laplacian smoothing.}
+	\label{tbl:estimation_rates}
+\end{table}
+
+\begin{table}
+	\begin{center}
+		\begin{tabular}{p{.1\textwidth} | p{.14\textwidth} p{.12\textwidth} }
+			Dimension & Laplacian smoothing \eqref{eqn:laplacian_smoothing_test} & Thin-plate splines \eqref{eqn:thin_plate_spline_test} \\
+			\hline
+			$d = 1$ & ${\bf n^{-4/5}}$ & \textcolor{red}{${\bf n^{-4/5}}$} \\
+			$d = 2,3$ & ${\bf n^{-4/(4 + d)}}$ & \textcolor{red}{$n^{-1/2}$} \\
+			$d \geq 4$\footnotemark & ${ \bf n^{-1/2}}$ & \textcolor{red}{${ \bf n^{-1/2}}$}
+		\end{tabular}
+	\end{center}
+	\caption{Summary of testing rates over fixed radius first-order Sobolev balls; black, red, and bold font used as in Table~\ref{tbl:estimation_rates}.
+Rates for $d \geq 4$ assume $f_0 \in \Leb^4(\Xset,M)$.
+When $\Xset$ is an $m$ dimensional manifold embedded in $\Rd$, all rates hold with $d$ replaced by $m$.}
+	\label{tbl:testing_rates}
+\end{table}
+\footnotetext{Assuming $f_0 \in \Leb^4(\Xset,M)$.}
+\paragraph{Future extensions.}
+To be clear, these results are not the end of the story.
+For one thing, the estimator $\wh{f}$ is defined only at the design points $\{X_1,\ldots,X_n\}$.
+As such, we measure loss using the in-sample mean squared error
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 := \frac{1}{n}\sum_{i = 1}^{n} \Bigl(\wh{f}_i - f_0(X_i)\Bigr)^2.
+\end{equation*}
+In Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we briefly discuss how one could extend $\wh{f}$ to a function over all $\Xset$, in such a way that the out-of-sample mean square error $\|\wh{f} - f_0\|_{\Leb^2(\Xset)}$ remains small, but leave a formal analysis of the problem to future work.
+
+In a different direction, the problem~\eqref{eqn:thin_plate_spline} is only a special, first-order case of thin-plate splines. In general, the degree-$k$ thin plate spline is defined as a minimizer of
+\begin{equation*}
+\sum_{i = 1}^{n} (Y_i - f(X_i))^2 + \rho \cdot \sum_{|m| = k} \int_{\Xset} \bigl(D^mf(x)\bigr)^2 \,dx
+\end{equation*}
+where for each multiindex $m$ we write $D^m$ for the partial derivative $D^mf(x):= \partial^kf/\partial x_{1}^{m_1}\ldots \partial x_{d}^{m_d}$.
+Assuming the $k$th-order partial derivatives $D^mf_0$ are all bounded in $\Leb^2(\Xset)$ norm, the degree-$k$ thin plate spline has error on the order of $n^{-2k/(2k + d)}$ \citep{vandergeer2000}, which is the minimax optimal rate for such functions.
+Of course, assuming $f_0$ possesses $k$ bounded derivatives for some $2k > d$ is a rather strong assumption which may not be warranted, but at present we do not know whether (adaptations of) Laplacian smoothing on neighborhood graphs can achieve these faster rates.
+
+% \paragraph{Notation.}
+% AJG: Need to make sure these are all defined somewhere
+% For an integer $p > 0$, we use $\Leb^p(\Xset)$ to refer to the set of functions $f$ for which the $p$th moment $\norm{f}_{\Leb^p(\Xset)}^p := \int_{\Xset} |f(x)|^p \,dx < \infty$, and $C^p(\Xset)$ to refer to those function which are $p$-times continuously differentiable with bounded $p$th derivative.
+Finally, we use $C,C_1,\ldots$ and $c,c_1,\ldots$ as constants which are independent of $n$ and $f_0$, and for convenience, we will also use asymptotic notation: $a_n \lesssim b_n$ means that $a_n \leq Cb_n$ for some $C > 0$, and $a_n \asymp b_n$ means that $a_n \lesssim b_n$ and $b_n \lesssim a_n$.
+
+\section{BACKGROUND}
+\label{sec:problem_setup_and_background}
+Before we formally state our main results in Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we define neighborhood graph Laplacians, and review the known minimax rates over first-order Sobolev spaces.
+
+\paragraph{Neighborhood graph Laplacians.}
+As mentioned, in the graph-based approach to nonparametric regression we first build a neighborhood graph $G_{n,r} = ([n],W)$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense.
+The $n \times n$ weight matrix $W = (W_{ij})$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $W_{ij}$ are given by
 \begin{equation*}
 \label{eqn:neighborhood_graph}
-	{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr).
+W_{ij} = K\Biggl(\frac{\norm{X_i - X_j}}{r}\Biggr),
 \end{equation*}
-	Then the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$, and the graph Laplacian can be written as
+where $\norm{\cdot}$ is the Euclidean norm in $\Rd$.
+Then the degree matrix $D$ is the $n \times n$ diagonal matrix with entries $D_{ii} = \sum_{j = 1}^{n}W_{ij}$, and the graph Laplacian can be written as
 \begin{equation}
 \label{eqn:graph_Laplacian}
-	\Lap_{n,r} = \bf{D} - \bf{W}
+\Lap_{n,r} = D - W;
 \end{equation}
-	\item The Laplacian smoothing estimator $\wh{f}_{\LS}$ \citep{smola2003} is a penalized least squares estimator, given by
+we add the subscripts $n$ and $r$ to $\Lap$ and $G$ to emphasize that they depend on the random design points $X_1,\ldots,X_n$ and the radius $r$.
+To avoid any confusion: henceforth when speak of $\wh{f}$ and $\wh{T}$, we will always assume they are fit using the neighborhood graph Laplacian $L = L_{n,r}$.
+
+Note that $L$ is a symmetric, positive semi-definite matrix, and so admits an eigendecomposition $L = \sum_{k = 1}^{n} \lambda_k v_k v_k^T$.
+We order the eigenvalues $0 = \lambda_1 \leq \lambda_2 \leq \cdots \leq \lambda_n$, and always consider unit-norm eigenvectors $\|v_k\|_2^2 = 1$; we will make reference to these objects later on.
+
+\paragraph{Sobolev spaces.}
+Now that we've defined the neighborhood graph Laplacian $\Lap_{n,r}$, we step away from graph-based methods for a moment, to briefly recall some classical results regarding minimax rates over Sobolev classes.
+
+We say that a function $f \in \Leb^2(\Xset)$ belongs to the \emph{first-order Sobolev space} $H^1(\Xset)$ iff for each $j = 1,\ldots,d$, the weak partial derivative $D^jf$ exists and belong to $\Leb^2(\Xset)$.
+For such functions $f \in H^1(\Xset)$, the Sobolev seminorm $\seminorm{f}_{H^{1}(\Xset)}$ is the average size of the gradient;
+\begin{equation*}
+\seminorm{f}_{H^1(\Xset)}^2 = \int_{\Xset} \bigl\|\nabla f(x)\bigr\|^2 \,dx
+\end{equation*}
+with corresponding Sobolev norm $\norm{f}_{H^1(\Xset)} := \norm{f}_{\Leb^2(\Xset)} + |f|_{H^1(\Xset)}$. The Sobolev balls $H^1(\Xset,M)$ for $M > 0$ are then defined as
+\begin{equation*}
+H^1(\Xset,M) := \Bigl\{f \in H^1(\Xset): \norm{f}_{H^1(\Xset)}^2 \leq M^2\Bigr\}.
+\end{equation*}
+For further details regarding Sobolev spaces see e.g.~\citep{evans10,leoni2017}.
+
+\paragraph{Minimax rates.}
+To carry out a minimax analysis of regression in Sobolev spaces, one must impose some regularity conditions on the design distribution $P$. We shall assume the following.
+\begin{enumerate}[label=(P\arabic*)]
+	\item
+	\label{asmp:domain}
+	$P$ is supported on a domain $\Xset \subset \Rd$, an open, connected set with Lipschitz boundary.
+	\item
+	\label{asmp:bounded_lipschitz_density} 
+	$P$ admits a density $p$ bounded away from $0$ and $\infty$, i.e.
+	\begin{equation*}
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty,~~\textrm{for all $x \in \Xset$.}
+	\end{equation*}
+	Additionally, the restriction of $p$ to $\Xset$ is Lipschitz, with Lipschitz constant $L_p$.
+\end{enumerate}
+
+Under~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density}, the minimax estimation rate over Sobolev balls is (see e.g. \citep{tsybakov2008_book}),
 \begin{equation}
-	\label{eqn:laplacian_smoothing}
-	\wh{f}_{\LS} := \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_{n,r} f \biggr\}
+\label{eqn:sobolev_space_estimation_minimax_rate}
+\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{2d/(2 + d)}n^{-2/(2 + d)}.
 \end{equation}
-	where $\rho > 0$ acts a tuning parameter on the penalty $f^T \Lap_{{n,r}} f$.
-	Assuming~\eqref{eqn:laplacian_smoothing} is a reasonable estimator of $f_0$, the squared empirical norm
+
+As nonparametric hypothesis testing is (comparatively) less familiar than nonparametric estimation, we briefly summarize the main idea before stating the optimal error rate.
+In the goodness-of-fit testing problem, we ask for a test function---formally, a Borel measurable function $\phi$ that takes values in $\{0,1\}$--- which can distinguish between the hypotheses
 \begin{equation}
-	T_{\LS} = \frac{1}{n}\Bigl\|\wt{f}_{\LS}\Bigr\|_2^2 \label{eqn:laplacian_smoothing_test}
+\mathbf{H}_0: f_0 = f_0^{\star}, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{f_0^{\star}\}.
 \end{equation} 
-	is in turn a reasonable statistic to assess whether or not $f_0 = 0$. 
+Typically, the null hypothesis $f_0 = f_0^{\star} \in \mc{F}$ reflects the absence of interesting structure, and $\mc{F} \setminus  \{f_0^{\star}\}$ is a set of smooth departures from this null.
+To fix ideas, as in \cite{ingster2009} we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\star} = 0$ and $\mc{F} = H^1(\Xset,M)$ is a first-order Sobolev ball; this is without loss of generality.
+
+The minimax critical radius is the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{F}_{\epsilon} := \mc{F} \cap \{f: \|f\|_{\Leb^2(\Xset)} \geq \epsilon\}$:
+\begin{equation*}
+\epsilon(\mc{F}) := \inf\Biggl\{\epsilon > 0: \inf_{\phi} \biggl[ \sup_{f_0 \in \mc{F}_{\epsilon}} \Ebb_{f_0}[1 - \phi]\biggr] \leq \alpha\Biggr\}
+\end{equation*} 
+where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation under the regression function $f_0$.\footnote{ Clearly, the minimax critical radius $\epsilon$ depends on $\alpha$.
+However we adopt the typical convention of treating $\alpha$ as a small but fixed positive constant; hence it will not affect the testing error rates, and we suppress it notationally.} See \citep{ingster82, ingster87, ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
+
+Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing critical radius over $H^1(\Xset,M)$ is much smaller than the minimax estimation rate \citep{ingster2009}:
+\begin{equation}
+\label{eqn:sobolev_space_testing_critical_radius}
+\epsilon^2\Bigl(H^1(\Xset,M)\Bigr) \asymp M^{2d/(4 + d)}n^{-4/(4 + d)}~~\textrm{for $1 \leq d < 4$.}
+\end{equation}
+When $d \geq 4$ the functions in $H^1(\Xset)$ are very irregular---formally speaking $H^1(\Xset)$ does not continuously embed into $\Leb^4(\Xset)$ when $d \geq 4$---and the minimax testing rates in this regime are unknown.
+
+\section{MINIMAX OPTIMALITY OF LAPLACIAN SMOOTHING}
+\label{sec:minimax_optimal_laplacian_smoothing}
+
+We now formalize the main conclusions of this work: that Laplacian smoothing methods on neighborhood graphs are minimax rate-optimal over first-order continuum Sobolev classes.
+Throughout this section, we use $C,C_1,\ldots$ and $c,c_1,\ldots$ to refer to constants which do not depend on $n$ or $f_0$.
+We will assume~\ref{asmp:domain} and ~\ref{asmp:bounded_lipschitz_density}, and additionally that the kernel $K$ satisfies condition~\ref{asmp:kernel}.
+\begin{enumerate}[label=(K\arabic*)]
+	\item
+	\label{asmp:kernel}
+	$K:[0,\infty) \to [0,\infty)$ is a non-increasing function supported on $[0,1]$, its restriction to $[0,1]$ is Lipschitz, and $K(1) > 0$. Additionally, it is normalized so that
+	\begin{equation*}
+	\int_{\Reals^d} K\bigl(\norm{z}\bigr) \,dz = 1.
+	\end{equation*}
+	Let $\sigma_K := \frac{1}{d} \int_{\Rd} \|x\|^2 K(\|x\|) \,dx$.
+\end{enumerate}
+This is a relatively mild condition: the choice of kernel is under the user's control, and moreover~\ref{asmp:kernel} covers many (though not all) common kernel choices.
+
+\paragraph{Estimation error of Laplacian smoothing.} 
+Under these conditions, the Laplacian smoothing estimator $\wh{f}$ achieves, up to constants, the minimax risk $n^{-2/(2 + d)}$ over $H^1(\Xset,M)$.
+This statement will hold whenever the graph $G_{n,r}$ is computed with radius $r$ in the following range.
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{0}
+	\item 
+	\label{asmp:ls_kernel_radius_estimation}
+	The neighborhood graph radius $r$ satisfies
+	\begin{equation*}
+	C_0\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq r \leq n^{-3/(4 + 2d)} M^{(d - 4)/(4 + 2d)} \wedge c_0.
+	\end{equation*}
+\end{enumerate}
+We comment on~\ref{asmp:ls_kernel_radius_estimation} after stating Theorem~\ref{thm:laplacian_smoothing_estimation1}.
+The proof of this theorem, as with the proofs of all results in this paper, can be found in the supplementary document.
+\begin{theorem}
+	\label{thm:laplacian_smoothing_estimation1}
+	Let $f_0 \in H^1(\Xset,M)$ with $d < 4$.
+Suppose that that the neighborhood graph $G_{n,r}$ is computed with a radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_estimation},  and the Laplacian smoothing estimator $\wh{f}$ with $\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$.
+With probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c M^{d/(2d + 4)} n^{d/(2+d)})$, it holds that
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_estimation1}
+	\Bigl\|\wh{f} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(2 + d)} n^{-2/(2 + d)}.
+	\end{equation}
+\end{theorem}
+To summarize: when $d = 1,2$ or $3$, with high probability the Laplacian smoothing estimator $\wh{f}$ has in-sample mean squared error within a constant factor of the minimax risk. Some remarks:
+\begin{itemize}
+	\item The lower bound on $r$ imposed by~\ref{asmp:ls_kernel_radius_estimation} is  compatible with practice, where by far the most common choice of radius is the connectivity threshold $r \asymp (\log(n)/n)^{1/d}$, chosen to make the graph $G_{n,r}$ as sparse as possible while still being connected, for maximum computational efficiency.
+The upper bound is a bit more mysterious.
+We need it for technical reasons to ensure that $\wh{f}$ does not overfit, but note that as a practical matter one rarely chooses $r$ to be so large.
+	\item It is possible to extend $\wh{f}$ to be defined on all of $\Xset$, and evaluate the error of such an extension in $\Leb^2(\Xset)$ norm.
+Assuming that the extension of $\wh{f}$ and the regression function $f_0$ are both suitably smooth, tools from empirical process theory (Chapter 14 of \cite{wainwright2019}) or approximation theory (Section 15.5 of \citep{johnstone2011}) should guarantee that the $\Leb^2(\Xset)$ error is not too much greater than its in-sample counterpart.
+In fact, as we show in our supplementary material, if $f_0$ is Lipschitz smooth and we extend $\wh{f}$ to be piecewise constant over the Voronoi tesselation induced by $X_1,\ldots,X_n$, the out-of-sample error $\|f_0 - \wh{f}\|_{\Leb^2(X)}$ is within a negligible factor of the in-sample error $\|f_0 - \wh{f}\|_n$, but we leave analysis of the Sobolev case to future work.
+	\item Under the stronger assumption that $f_0$ is Lipschitz smooth, we can replace the factor of $\delta$ in the high probability bound by the sharper $\delta^2/n$.
 \end{itemize}
 
-\textbf{(2.a)} Laplacian smoothing has many advantages compared to other nonparametric regression methods. For instance:
+When $d = 4$, by setting $r \asymp (\log n/n)^{1/4}$ our analysis results in an upper bound on the mean squared error within a $(\log n)^{1/3}$ factor of the minimax risk, but when $d \geq 5$ our analysis yields upper bounds that are much worse than the minimax rate.
+This mirrors the conclusions of~\cite{sadhanala16}, who investigate estimation rates of Laplacian smoothing over the $d$-dimensional grid graph.
+\cite{sadhanala16} argue that their analysis is tight, and that it is the estimator, rather than the analysis, which is deficient when $d \geq 5$.
+Formally proving such a claim turns out to be harder in the random design setting than in the fixed design setting, and we leave it for future work.
 
+However, we do investigate the matter empirically.
+In Figure~\ref{fig:fig1}, we consider the mean squared error of the Laplacian smoothing estimator, for several different values of dimension $d$.
+Here, $X_1,\ldots,X_n$ are sampled uniformly over $\Xset = [-1,1]^d$, and the regression function $f_0(x) \propto \Pi_{i = 1}^{d} \cos(m \pi x_i)$, where $m = 2$ for $d = 2$, and $m = 1$ for $d \geq 3$.
+This is a rather smooth regression function, and when $d = 2$ and $d = 3$ empirically Laplacian smoothing appears to achieve or exceed the minimax rate.
+When $d = 4$ Laplacian smoothing appears modestly suboptimal; this fits with our theoretical upper bound, which includes a $(\log n)^{1/3}$ factor that plays a non-negligible role for these problem sizes.
+On the other hand, when $d = 5$ Laplacian smoothing seems to be decidedly suboptimal.
+
+\begin{figure}[thb]
+	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_2d.pdf}
+	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_3d.pdf}\\ 
+	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_4d.pdf}
+	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_5d.pdf}
+
+	\caption{Mean squared error of Laplacian smoothing (\texttt{LS}) as a function of sample size $n$.
+Plots are on the log-log scale, MSE is averaged over 5 iterations, with Laplacian smoothing tuned for optimal MSE.
+The black line shows the minimax rate.}
+	\label{fig:fig1}
+\end{figure}
+
+\paragraph{Testing error of Laplacian smoothing.}
+Let us define a test using the statistic $\wh{T}$. For $b \geq 1$, set the threshold $\wh{t}_b$ as
+\begin{equation*}
+\wh{t}_{b} := \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k + 1\bigr)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k + 1\bigr)^4}},
+\end{equation*}
+where we recall $\lambda_k$ is the $k$th smallest eigenvalue of $\Lap_{n,r}$; the Laplacian smoothing test is simply $\wh{\varphi} := \1\bigl\{\wh{T} \leq \wh{t}_b\bigr\}$.
+(Clearly the test $\wh{\varphi}$ depends on $b$, but we suppress this notationally.) The number $b \geq 1$ is a tuning parameter selected by the user, with the choice made based on the tolerated level of Type I and Type II error.
+In fact, the threshold $\wh{t}_b$ is precisely the right choice to control the Type I error of $\wh{\varphi}$; as we show in our supplementary material,
+\begin{equation}
+\label{eqn:type_I_error}
+\Ebb_0\bigl[\wh{\varphi}\bigr] \leq \frac{1}{b^2}.
+\end{equation}
+Theorem~\ref{thm:laplacian_smoothing_testing} establishes that this test also has small Type II error, uniformly over all $f_0$ separated from $0$ by at least the critical radius $\epsilon\bigl(H^1(\Xset,M)\bigr)$ given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
+For this to hold, we will require a tight range of scalings for $r$.
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{1}
+	\item 
+	\label{asmp:ls_kernel_radius_testing}
+	The neighborhood graph radius $r$ satisfies
+	\begin{equation*}
+	C_0\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq r \leq M^{\frac{(d - 8)}{8 + 2d}}n^{\frac{d - 20}{32 + 8d}} \wedge c_0.
+	\end{equation*}
+\end{enumerate}
+Note that when $d < 4$, if $M$ is not too large there always exists a choice of $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}.
+\begin{theorem}
+	\label{thm:laplacian_smoothing_testing}
+	Let $f_0 \in H^1(\Xset,M)$ for $M \leq n^{(4 - d)/(4 + d)}$ and $d < 4$.
+Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}, and the Laplacian smoothing test $\wh{\varphi}$ with $\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$, and $b \geq 1$.
+For all $f_0$ such that
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_testing}
+	\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2d/(4 + d)} n^{-4/(4 + d)}
+	\end{equation} 
+	the Type II error is upper bounded:
+	\begin{align*}
+	& \Ebb_{f_0}\Bigl[1 - \wh{\varphi} \Bigr] \leq \\
+	& C\biggl(\frac{1}{b}\Bigl[1 + n^{-d(4 + d)}M^{-2d/(4 + d)}\Bigr] + C_1n\exp\bigl(-c_1nr^d\bigr)\biggr).
+	\end{align*}
+\end{theorem}
+A couple remarks:
 \begin{itemize}
-	\item It is fast, easy, and stable to compute.
-Letting ${\bf Y} = (Y_1,\ldots,Y_n) \in \Reals^n$, it is not hard to see that the minimizer $\wt{f}_{\LS} = (\rho \Lap_{n,r} + I)^{-1}{\bf Y}$ can be computed by solving a linear equation; in practice, one typically chooses $r$ to be small, with the result being that the matrix $\Lap_{n,r}$ is sparse.
-There exist known fast solvers of exactly this system.
-	\item Is is generalizable to non-standard data---for instance, text or images, or really any data modality on which it is possible to define a kernel.
-	\item It is easily adaptable to the semi-supervised learning setting. 
+	\item As mentioned in Section~\ref{sec:problem_setup_and_background}, when $d \geq 4$ the Sobolev balls $H^1(\Xset,M)$ include quite irregular functions $f \not\in \Leb^4(\Xset)$.
+Proving tight lower bounds over such classes is non-trivial, and to the best of our knowledge such an analysis remains outstanding.
+On the other hand, if we explicitly assume that $f_0 \in \Leb^4(\Xset,M)$, then \cite{guerre02} show that the testing problem is characterized by the dimension-free lower bound $\epsilon^{2}(\Leb^4(\Xset,M)) \gtrsim n^{-1/2}$.
+Moreover, by training $\wh{f}$ to the interpolation limit, i.e.
+setting $\rho = 0$, the test $\wh{\varphi}$ achieves (up to constants) this lower bound.
+That is, for any $f_0 \in \Leb^4(\Xset,M)$ such that $\|f_0\|_{\Leb^2(\Xset)}^2 \geq C b^2n^{-1/2}$, we have both
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_testing_low_smoothness}
+	E_{f_0}\Bigl[1 - \wh{\varphi}\Bigr] \leq \frac{C(1 + M^4)}{b^2}
+	\end{equation} 
+	and $\Ebb_0[\wh{\varphi}] \leq 1/b^2$. These same results apply to $\wt{T}$, since when $d > 1$ the estimator $\wt{f}$ will interpolate the responses $Y_1,\ldots,Y_n$.
+	\item To compute the data-dependent threshold $\wh{t}_b$, one must know each of the eigenvalues $\lambda_1,\ldots,\lambda_n$.
+Computing all $n$ of these eigenvalues is far more expensive---on the order of $n^3$ operations---than computing $\wh{T}$.
+That being said, in practice we would not recommend using $\wh{t}_b$ anyway, and would instead we make the standard recommendation to calibrate via a permutation test \citep{hoeffding1952}.
 \end{itemize}
 
-\textbf{(2.b)} For this reason a considerable body of work has emerged analyzing the consistency of the graph Laplacian $\Lap_{{n,r}}$ as $n \to \infty$ and $r(n) \to 0$.
-This is meant in various senses:
-\begin{itemize}
-	\item Pointwise consistency, meaning $\Lap_{n,r}f \to \Delta_Pf$ for an appropriate limiting operator $\Delta_P$.
-	\item Spectral consistency, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$.
-	\item Consistency of norms, meaning $f^T \Lap_{n,f} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$. 
-\end{itemize}
-However, the statistical optimality of Laplacian smoothing is still not well understood.
+\paragraph{Comparison with thin-plate splines.}
+With some of our main results in hand, let us pause to offer some explanation of why Laplacian smoothing can be optimal in settings where thin-plate splines are not even consistent.
 
-\textbf{(2.c)} Our main contributions lie in this latter direction, and they can all be summarized as follows: Laplacian smoothing methods are minimax optimal over Sobolev spaces.
-In more detail, we will show that when $f_0$ belongs to the Sobolev ball $H^1(\Xset;M)$:
-\begin{itemize}
-	\item \textbf{Minimax optimal estimation.} With high probability, $\frac{1}{n}\norm{\wt{f}_{\LS} - f_0}_{2}^2 \lesssim n^{-2/(2 + d)}$.
-	\item \textbf{Minimax optimal testing.}
-	A level-$\alpha$ test constructed using $T_{\LS}$ has non-trivial power whenever $\norm{f_0}_{\Leb^2(\Xset)}^2 \gtrsim n^{-4/(4 + d)}$. 
-	\item \textbf{Manifold adaptivity.}
-	If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$, both of the aforementioned rates hold with $d$ replaced by $m$.
-\end{itemize}
-Finally, we will also establish that the Laplacian smoothing estimator can be altered to take advantage of higher-order smoothness assumptions on $f_0$.
-This last conclusion will hold only under very limited circumstances.
+First, we should make clear why this difference in performance is so surprising.
+As mentioned previously, the penalties in~\eqref{eqn:laplacian_smoothing} and~\eqref{eqn:thin_plate_spline} can be closely tied together: \cite{bousquet03} show that for any $f \in C^2(\Xset)$,
+\begin{equation}
+\label{eqn:seminorm_consistency}
+\begin{aligned}
+\lim \frac{1}{n^2 r^{d + 2}} f^T \Lap_{n,r} f & = \int_{\Xset} f(x) \cdot \Delta_Pf(x) p(x) \,dx \\
+& = \int_{\Xset} \|\nabla f(x)\|^2 p^2(x) \,dx
+\end{aligned}
+\end{equation}
+in the above, the limit is as $n \to \infty$ and $r \to 0$, $\Delta_P$ is the (weighted) Laplace-Beltrami operator
+\begin{equation*}
+\Delta_Pf(x) := -\frac{1}{p} \dive\bigl(p^2\nabla f(x))
+\end{equation*}
+and the second equality follows from integration by parts.\footnote{Assuming $f$ satisfies Dirichlet boundary conditions.} To be clear, this argument does not formally imply that $\wh{f}$ and $\wt{f}$ are close---note that~\eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:thin_plate_spline} is $H^1(\Xset)$---but it does seem to suggest that the two estimators should behave somewhat similarly.
 
-\paragraph{Organization.}
+Of course, we know this is not the case: $\wh{f}$ and $\wt{f}$ look very different when $d > 1$.
+What is driving this difference?
+The key point is that the discretization imposed by the graph $G_{n,r}$---which seemed so problematic at first glance---turns out to be a blessing.
+The problem with~\eqref{eqn:thin_plate_spline} is that the function class $H^1(\Xset)$ is far too large when $d \geq 2$.
+This is meant in various related senses.
+From the Sobolev embedding theorem \cite{evans10}, we know that $H^1(\Xset)$ does not continuously embed into any H\"{o}lder space, and that $H^1(\Xset)$ is not an Reproducing Kernel Hilbert Space.
+More relevant to this context, by placing ``bump functions'' of arbitrarily small radius around each of the design points $X_1,\ldots,X_n$, one can interpolate the responses $Y_1,\ldots,Y_n$ in such a manner that forces the criterion in~\eqref{eqn:thin_plate_spline} to 0 \citep{green93}; the resulting estimator will clearly be inconsistent.
 
-List out what is to come.
+On the other hand~\eqref{eqn:laplacian_smoothing} is a finite dimensional (if growing linearly with $n$) problem---as a result $\wh{f}$ has far less capacity to overfit than does $\wt{f}$.
+Of course, discretization is not the only way to make the problem~\eqref{eqn:thin_plate_spline} more tractable: for instance, one could replace the penalty $\int_{\Xset} \|\nabla f(x)\|^2 \,dx$ by the stricter choice $\sup_{x \in \Xset} \|\nabla f(x)\|$,  or conduct the problem over some finite dimensional linear subspace of $H^1(\Xset)$ (i.e use a sieve).
+These solutions improve the statistical properties of $\wt{f}$ when $d > 1$ \citep{birge1993,birge1998,vandergeer2000}, but the first is computationally intractable, while the second requires the estimator to be specifically tailored to the domain $\Xset$, in stark contrast to $\wh{f}$.
 
-\section{Minimax regression over Sobolev spaces}
+\paragraph{Overview of analysis.}
+The comparison with thin-plate splines highlights some surprising strengths of $\wh{f}$.
+Unfortunately, this comparison also precludes analyzing $\wh{f}$ by using~\eqref{eqn:seminorm_consistency} to establish a coupling between $\wh{f}$ and $\wt{f}$; we know this won't work, because we would like to prove tight error bounds on $\wh{f}$ in regimes where no such bounds exist for $\wt{f}$.
 
-\textbf{(3)} Sobolev spaces are classical function spaces, over which linear nonparametric regression methods.
-One notable such method is smoothing splines, to which we give special attention because of their close connection to Laplacian smoothing.
+Instead we take a different approach, and directly analyze the error of $\wh{f}$ and $\wh{T}$ using a (conditional on $X_1,\ldots,X_n$) bias-variance decomposition. A standard calculation shows that
+\begin{equation*}
+\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \underbrace{\frac{2\rho}{n} \bigl(f_0^T \Lap_{n,r} f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{10}{n} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2}}_{\textrm{variance}}
+\end{equation*}
+and likewise that $\wh{\varphi}$ has high power whenever
+\begin{equation*}
+\bigl\|f_0\bigr\|_n^2 \geq \underbrace{\frac{2\rho}{n} \bigl(f_0^T \Lap_{n,r} f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{4b}{n} \sqrt{\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}}}_{\textrm{variance}}
+\end{equation*}
+Note that the bias and variance terms are each functionals of the random graph $G_{n,r}$, and so are themselves random.
+To upper bound them, we build on some recent works \citep{burago2014,trillos2019,calder2019} regarding the consistency of neighborhood graphs to establish the following two Lemmas.
+In the second of these Lemmas, we will assume
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{2}
+	\item
+	\label{asmp:radius1}
+	The radius $r$ satisfies $C_0(\log n/n)^{1/d} \leq r \leq c_0$.
+\end{enumerate}
+and put $A_{n,r}(k) := \min\{nr^{d + 2}k^{2/d},nr^d\}$.
 
-\textbf{(4)} In fact, an alternative way to motivate Laplacian smoothing nonparametric regression methods---as opposed to the advantages mentioned in \textbf{(2.a)}---comes from viewing them as discrete-time versions of smoothing splines.
-This is made more rigorous by the consistency results of \textbf{(2.b)}.
-Viewed in this light, an additional advantage of Laplacian smoothing emerges: it automatically adapts to the geometry of $\Xset$ and $p$.
+\begin{lemma}
+	\label{lem:graph_sobolev_seminorm}
+	For any $f \in H^1(\Xset)$, with probability at least $1 - \delta$, it holds that
+	\begin{equation}
+	\label{eqn:graph_sobolev_seminorm}
+	f^T \Lap_{n,r} f \leq \frac{(1 + L_p)^2 \sigma_K}{\delta} n^2 r^{d + 2} |f|_{H^1(X)}^2.
+	\end{equation}
+\end{lemma}
+\begin{lemma}
+	\label{lem:neighborhood_eigenvalue}
+	For any neighborhood graph radius $r$ which satisfies~\ref{asmp:radius1}, with probability at least $1 - C_1n\exp(-c_1nr^d)$ it holds that
+	\begin{equation}
+	\label{eqn:neighborhood_eigenvalue}
+	c_2A_{n,r}(k) \leq \lambda_k \leq C_2A_{n,r}(k)~~\textrm{for all $2 \leq k \leq n$}
+	\end{equation}
+\end{lemma}
 
-\textbf{(5)} Smoothing splines are known to have strong minimax optimal properties. Reasoning by analogy, we might hope that Laplacian smoothing would share these properties.
+Lemma~\ref{lem:graph_sobolev_seminorm} gives a direct upper bound on the bias term.
+Lemma~\ref{lem:neighborhood_eigenvalue} results in a sufficient upper bound on the variance term whenever the neighborhood graph radius $r$ is \textit{sufficiently small}; precisely, when $r$ is upper bounded as in~\ref{asmp:ls_kernel_radius_estimation} (for estimation) or~\ref{asmp:ls_kernel_radius_testing} (for testing).
+The parameter $\rho$ is then tuned to minimize the sum of bias and variance, in the usual manner.
 
-\section{Graph Sobolev classes}
+It is useful to give another perspective on our approach.
+When analyzing penalized least squares estimators such as $\wh{f}$ or $\wt{f}$ one typically assumes two properties: first, that the regression function $f_0$ lies in (or near) a ball in the normed space induced by the penalty, and second that this ball is reasonably small, e.g.
+as measured by metric entropy.
+In contrast, the Laplacian smoothing penalty induces a ball
+\begin{equation*}
+H^1(G_{n,r},M) := \{f: f^T \Lap_{n,r} f \leq M^2\}
+\end{equation*}
+that is data-dependent and random, and so we do not have access to either of the aforementioned properties \emph{a priori}; instead we must prove they hold with high probability.
+In this sense, our analysis is rather different than is typical in nonparametric regression.
 
-\textbf{(6)} The minimax optimality of smoothing splines rests on two crucial facts: one, the a priori assumption that $|f|_{H^1(\Xset)}$ is bounded; two, an upper bound on the metric entropy of $H^1(\Xset)$, characterized by the decay of its eigenvalues.
-\textcolor{red}{(TODO)}:(I know this isn't well-stated, and I need to clean it up.)  We now show that qualitatively similar statements hold with respect to
+\section{MANIFOLD ADAPTIVITY}
+\label{sec:manifold_adaptivity}
+The minimax rates $n^{-2/(2 + d)}$ (in estimation) and $n^{-4/(4 + d)}$ (in testing) suffer from the curse of dimensionality.
+However, in practice it is often reasonable to assume a \emph{manifold hypothesis}: roughly speaking, that the data $X_1,\ldots,X_n$ lie on a submanifold $\Xset$ of $\Rd$ that has intrinsic dimension $1 \leq m \leq d$.
+Under such an assumption, it is known \citep{bickel2007, ariascastro2018} that the optimal rates over $H^1(\Xset)$ now scale like $n^{-2/(2 + m)}$ (for estimation) and $n^{-4/(4 + m)}$ (for testing), which can be much smaller than the full-dimensional error rates when $m \ll d$.
 
-\quad \textbf{(6.1)} the graph Sobolev semi-norm $f_0^T \Lap_{n,r} f_0$ and,
+On the other hand, a theory has been developed~\citep{belkin03,belkin05,niyogi2013} establishing the the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+In this section, we contribute to this line of work by showing that under the manifold hypothesis, Laplacian smoothing achieves the sharper minimax rates over $H^1(\Xset)$.
 
-\quad \textbf{(6.2)} the graph eigenvalues $\lambda_k(\Lap_{n,r})$.
+\paragraph{Error rates with the manifold hypothesis.}
+The conditions and results of this Section will be largely similar to the previous one, except with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
+For the rest of this section we assume the following.
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{2}
+	\item 
+	\label{asmp:domain_manifold}
+	The measure $P$ is supported on a compact, connected, smooth submanifold $\Xset$ of $\Reals^d$.
+The manifold $\Xset$ is of fixed dimension $1 \leq m \leq d$, is without boundary, and has positive reach \citep{federer1959}.
+	\item 
+	\label{asmp:density_manifold} $P$ has a density $p$ with respect to the volume form of $\Xset$ which is bounded away from $0$ and $\infty$,
+	\begin{equation*}
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty
+	\end{equation*}
+	for all $x \in \Xset$. The density $p$ is Lipschitz with Lipschitz constant $L_p$.
+\end{enumerate}
+Under the assumptions~\ref{asmp:domain_manifold}, \ref{asmp:density_manifold} and~\ref{asmp:kernel}, and for a suitable range of $r$, the error bounds on the estimator $\wh{f}$ and test $\wh{\varphi}$ will depend on $m$ instead of $d$.
 
-\section{Minimax optimality of Laplacian smoothing}
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{3}
+	\item 
+	\label{asmp:ls_kernel_radius_estimation_manifold}
+	The neighborhood graph radius $r$ satisfies
+	\begin{equation*}
+	C_3\biggl(\frac{\log n}{n}\biggr)^{1/m} \leq r \leq n^{\frac{-3}{(4 + 2m)}} M^{\frac{(m - 4)}{(4 + 2m)}} \wedge c_3.
+	\end{equation*}
+\end{enumerate}
+\begin{theorem}
+	\label{thm:laplacian_smoothing_estimation_manifold}
+	Let $f_0 \in H^1(\Xset,M)$ with $m < 4$.
+Suppose that that the neighborhood graph $G_{n,r}$ is computed with a radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_estimation_manifold},  and the Laplacian smoothing estimator $\wh{f}$ with $\rho = M^{-4/(2 + m)} (nr^{m + 2})^{-1} n^{-2/(2 + m)}$.
+With probability at least $1 - \delta -  C_4n\exp(-c_4nr^m) - \exp(-c M^{m/(2m + 4)} n^{m/(2+m)})$, it holds that
+	\begin{equation*}
+	\Bigl\|\wh{f} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2m/(2 + m)} n^{-2/(2 + m)}.
+	\end{equation*}
+\end{theorem}
 
-\textbf{(7)} Using \textbf{(6.1)} and \textbf{(6.2)}, we establish upper bounds on the estimation and testing error of Laplacian smoothing, which confirm that Laplacian smoothing is asymptotically minimax rate-optimal over the Sobolev class $H^1(\Xset)$.
+We require the following range of $r$ for testing.
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{4}
+	\item 
+	\label{asmp:ls_kernel_radius_testing_manifold}
+	The neighborhood graph radius $r$ satisfies
+	\begin{equation*}
+	C_3\biggl(\frac{\log n}{n}\biggr)^{1/m} \leq r \leq M^{\frac{(m - 8)}{8 + 2m}}n^{\frac{m - 20}{32 + 8m}} \wedge c_3.
+	\end{equation*}
+\end{enumerate}
 
-\textbf{(8)} While we motivated graph Laplacians by suggesting they might replicate the statistical properties of smoothing splines, we have in fact managed to show something stronger. In particular:
+\begin{theorem}
+	\label{thm:laplacian_smoothing_testing_manifold}
+	Let $f_0 \in H^1(\Xset,M)$ for $M \leq n^{(4 - m)/(4 + m)}$ and $m < 4$.
+Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing_manifold}, and the Laplacian smoothing test $\wh{\varphi}$ with $\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$, and $b \geq 1$.
+If
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_testing_manifold}
+	\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2m/(4 + m)} n^{-4/(4 + m)},
+	\end{equation} 
+	the Type II error is upper bounded:
+	\begin{align*}
+	& \Ebb_{f_0}\Bigl[1 - \wh{\varphi} \Bigr] \leq \\
+	& C\biggl(\frac{1}{b}\Bigl[1 + n^{-m(4 + m)}M^{-2m/(4 + m)}\Bigr] + C_4n\exp\bigl(-c_4nr^m\bigr)\biggr).
+	\end{align*}
+\end{theorem}
+The proof of Theorems~\ref{thm:laplacian_smoothing_estimation_manifold} and~\ref{thm:laplacian_smoothing_testing_manifold} proceeds in a similar manner to that of Theorems~\ref{thm:laplacian_smoothing_estimation1} and~\ref{thm:laplacian_smoothing_testing}.
+The key difference is that in the manifold setting, the equations~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} that we use to upper bound bias and variance will hold with $d$ replaced by $m$.
 
-\quad \textbf{(8.1)} When $d \geq 2$, there is no Sobolev embedding of $H^1(\Xset) \subseteq C^s(\Xset)$ for any $s > 0$. Therefore the smoothing spline estimator is ill-posed.
+We emphasize that very little about $\Xset$ need be known for Theorems~\ref{thm:laplacian_smoothing_estimation_manifold} and~\ref{thm:laplacian_smoothing_testing_manifold} to hold.
+Indeed, all that is needed is the intrinsic dimension $m$--- in order to properly tune $r$ and $\rho$--- and otherwise $\wh{f}$ and $\wh{\varphi}$ are computed without regard to $\Xset$.
+In contrast, the penalty in~\eqref{eqn:thin_plate_spline} would have to be specially tailored to work in this setting, revealing another advantage of Laplacian smoothing compared to thin-plate splines.
 
-\quad \textbf{(8.2)} Additionally, the metric entropy of $C^1(\Xset)$ is too large to obtain optimal rates for \textcolor{red}{Holder-smoothing splines} when $d \geq 2$.
-Thus, we see a gap emerge between the optimality properties of the discrete-time (Laplacian smoothing) and continuous-time (smoothing splines) estimators, in favor of the latter.
+\section{DISCUSSION}
+\label{sec:discussion}
+In this work, we have shown that Laplacian smoothing over a neighborhood graph is optimal for both estimation and goodness-of-fit testing problems.
 
-\quad \textbf{(8.3)} On the other hand, when $d \geq 4$, we no longer get the known optimal rates in the estimation problem.
-When $d = 4$, our rates are suboptimal by a factor of $\log n$.
-When $d > 5$ they are very suboptimal.
-We believe this is related to classical results regarding the entropy of low-smoothness Holder classes, i.e.
-the results used to justify \textbf{(8.2)}.
-The interesting point is that the parameter spaces $C^1(\Xset)$ and $H^1(G_{n,r})$ become ``too large'' for different values of $d$.
-
-\quad \textbf{(8.4)} \textcolor{red}{(TODO)}: Translate these results into error in $\Leb^2(P)$ error, if possible.
-
-\subsection{Manifold adaptivity}
-
-\quad \textbf{(9)} We now show that when $\Xset$ is an $m$-dimensional submanifold of $\Reals^d$, Laplacian smoothing approaches automatically ``feel'' the intrinsic dimension $m$ of $\Xset$, which translate into improved upper bounds on estimating and testing error.
-This confirms the statement in \textbf{(4)} that Laplacian smoothing automatically adapts to the geometry of $\Xset$ and $p$.
-
-\subsection{Higher-order smoothness classes}
-
-\quad \textbf{(10)} Analogously to smoothing splines, Laplacian smoothing can also be adapted to take advantage of additional assumed regularity on $f_0$, i.e.
-$f_0 \in H^s(\mc{\Xset})$ for $s > 1$.
-In very limited circumstances, we can show that this adapted estimator achieves the sharper minimax rates over these classes.
-However, the general story--for all combinations of $s$ and $d$---remains beyond our reach.
-
-\section{Simulations}
-
-\quad \textbf{(11)} Empirically, we demonstrate that the risk for Laplacian smoothing is comparable to that of smoothing splines, in the full dimensional case.
-Although one cannot carry out a comparison in the case of an unknown manifold --- since the smoothing spline estimator must be changed in a non-obvious manner --- we do show that when the density $p$ is concentrated close to a manifold and $f_0$ is smooth where $p$ is peaked, we have improved performance for $\wh{f}_{\LS}$ compared to $\wh{f}_{\mathrm{SM}}$.
-
-\section{Discussion}
-
-\clearpage
+There are still several extensions we feel are worth pursuing.
+In practice, it is more common to use $k$NN graphs than neighborhood graphs, due to the guaranteed connectivity and sparsity of the former; we suspect that by building on the work of~\cite{calder2019}, one can show that our main results all hold when the $k$NN graph is used.
+In another direction, one can also generalize Laplacian smoothing by replacing the penalty $f^T \Lap_{n,r} f$ with $f^T \Lap_{n,r}^k f$ for some $k > 1$.
+The hope is that this adapted procedure would be minimax optimal over the previously mentioned higher-order Sobolev classes $H^k(\Xset)$.
+In the very special case of $k = 2$ and $d \leq 2$, we can show that this is indeed the case, but the general story--for all combinations of $k$ and $d$---remains beyond our reach.
 

### 2020-10-20 notes/manuscripts/AISTATS/supplement_minimax_optimal_laplacian_smoothing.tex
Added proof of stronger results under Lipschitz assumptions.

@@
 \documentclass[twoside]{article}
-\usepackage{aistats2021}
-\usepackage{amsmath}
-\usepackage{amsfonts, amsthm, amssymb}
-\usepackage{graphicx}
-\usepackage[colorlinks]{hyperref}
+\usepackage{amsmath, amsfonts, amsthm, amssymb}
 \usepackage[parfill]{parskip}
@@
 \usepackage{tikz}
+\usepackage{verbatim}
+\usepackage{fullpage}
 
@@
 \newcommand{\wh}[1]{\widehat{#1}}
+\newcommand{\wc}[1]{\widecheck{#1}}
 \newcommand{\ol}[1]{\overline{#1}}
@@
 
+%%% So as not to overlap with theorem/lemma numbering from the main text...
+\setcounter{theorem}{4}
+\setcounter{lemma}{2}
+
 \theoremstyle{definition}
@@
 
+%%% Reference main text
+\usepackage{xr-hyper}
+\usepackage[colorlinks]{hyperref}
+\externaldocument{minimax_optimal_laplacian_smoothing}
+
+%%% New equation numbering
+\renewcommand{\theequation}{A.\arabic{equation}}
+
 \begin{document}
 
-\onecolumn
-\aistatstitle{Supplementary materials for: \\
-	Minimax Optimal Laplacian Smoothing}
-
-\section{Organization of Supplement}
-
-Our results all follow the same general two-part proof strategy, centered on conditioning.
-First, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to functionals of the graph $G$, and allow us to upper bound the error of $\wh{f}_{\LS}$ and $T_{\LS}$ conditional on $\mathbf{X} = x$.
-Second, we analyze the behavior of these functionals with respect to the particular neighborhood graph $G_{n,r}$ and give high probability (upper or lower) bounds.
-It is in this second step where we invoke our various assumptions on the distribution $P$ and regression function $f_0$.
+\title{Supplementary Materials for Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs}
+\date{}
+\maketitle
+
+In this supplement, we provide proofs of all results found in the paper ``Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs''.
+Our main theorems (Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing_manifold}) all follow the same general proof strategy, centered on conditioning.
+In Section~\ref{sec:fixed_graph_error_bounds}, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to (functionals of) the graph $G$, and allow us to upper bound the error of $\wh{f}$ and $\wh{\varphi}$ conditional on the design $\mathbf{X} = x$.
+In Sections~\ref{sec:graph_sobolev_seminorm}, \ref{sec:graph_eigenvalues}, and \ref{sec:empirical_norm}, we give all our necessary probabilistic estimates on these functionals, for the particular random neighborhood graph $G = G_{n,r}$.
+It is in these sections where we invoke our various assumptions on the distribution $P$ and regression function $f_0$.
+In Section~\ref{sec:main_results}, we prove our main theorems and some other results.
+In Section~\ref{sec:concentration}, we state a few concentration bounds that we use repeatedly in our proofs.
 
@@
 \label{sec:fixed_graph_error_bounds}
-
-Suppose we observe a fixed graph $G = \bigl([n],W\bigr)$ with Laplacian $\Lap_G$ and responses
+In this section, we adopt the fixed design perspective; or equivalently, condition on $X_i = x_i$ for $i = 1,\ldots,n$.
+Let $G = \bigl([n],W\bigr)$ be a fixed graph on $\{1,\ldots,n\}$ with Laplacian matrix $\Lap = D - W$.
+The randomness thus all comes from the responses
 \begin{equation}
 \label{eqn:fixed_graph_regression_model}
-Y_i = f_{0,i} + \varepsilon_i
-\end{equation}
-where $f_0 = (f_{0,1},\ldots,f_{0,n}) \in \Reals^n$, and the noise variables $\varepsilon_i$ are independent $N(0,1)$. The Laplacian smoothing estimator of $f_0$ on $G$ is
-\begin{equation}
+Y_i = f_{0}(x_i) + \varepsilon_i
+\end{equation}
+where the noise variables $\varepsilon_i$ are independent $N(0,1)$.
+In the rest of this section, we will mildly abuse notation and write $f_0 = (f_0(x_1),\ldots,f_0(x_n)) \in \Reals^n$.
+We will also write ${\bf Y} = (Y_1,\ldots,Y_n)$.
+
+Recall~\eqref{eqn:laplacian_smoothing} and~\eqref{eqn:laplacian_smoothing_test}: the Laplacian smoothing estimator of $f_0$ on $G$ is
+\begin{equation*}
 \label{eqn:ls_G}
-\wh{f}_{\LS}(G) := \argmin_{f \in \Reals^n} \biggl\{ \sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_G^{s}  f \biggr\} = (\rho \Lap_G^s + \Id)^{-1}Y.
-\end{equation}
-The statistic
-\begin{equation}
+\wh{f} := \argmin_{f \in \Reals^n} \biggl\{ \sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap f \biggr\} = (\rho \Lap + \Id)^{-1}{\bf Y}.
+\end{equation*}
+and the Laplacian smoothing test statistic is 
+\begin{equation*}
 \label{eqn:ls_ts_G}
-T_{\LS}(G) := \frac{1}{n} \Bigl\|\wh{f}_{\LS}(G)\Bigr\|_2^2 
-\end{equation}
-is used to distinguish
-\begin{equation*}
-\mathbf{H}_0: f_{0} = (0,...,0) ~~\textrm{versus}~~ \mathbf{H}_a: f_{0} \neq (0,...,0).
-\end{equation*}
-We note that~\eqref{eqn:laplacian_smoothing} and \eqref{eqn:laplacian_smoothing_test} specialize~\eqref{eqn:ls_G} and \eqref{eqn:ls_ts_G} to the case where $G = G_{n,r}$ and $s = 1$.
+\wh{T} := \frac{1}{n} \Bigl\|\wh{f}\Bigr\|_2^2.
+\end{equation*}
+We note that in this section, many of the derivations involved in upper bounding the estimation error of $\wh{f}$ are similar to those of \citep{sadhanala16}, with the difference being that we seek bounds in high probability rather than in expectation.
+We keep the work here self-contained for purposes of completeness.
 
 \subsection{Error bounds for linear smoothers}
-Let $S \in \Reals^{n \times n}$ be a square, symmetric matrix, and let $\check{f} = SY$ be a linear estimator of $f_0$.
-Clearly \eqref{eqn:ls_G} can be written in this form, and Lemma~\ref{lem:linear_smoother_fixed_graph_estimation} is therefore useful in analyzing it.
-Let $\lambdavec(S) = (\lambda_1(S),\ldots,\lambda_n(S)) \in \Reals^n$ denote the eigenvalues of $S$ and let $v_k(S)$ denote the corresponding \emph{unit-norm} eigenvectors, so that $S = \sum_{k = 1}^{n} \lambda_k(S) \cdot v_k(S) v_k(S)^T$.
-Denote $Z_k = v_k(S)^T \varepsilon$, and observe that $Z = (Z_1,\ldots,Z_n) \sim N(0,\Id)$.
+Let $S \in \Reals^{n \times n}$ be a square, symmetric matrix, and let 
+\begin{equation*}
+\wc{f} := SY
+\end{equation*}
+be a linear estimator of $f_0$.
+In  Lemma~\ref{lem:linear_smoother_fixed_graph_estimation} we upper bound the error $\frac{1}{n}\|\wc{f} - f_0\|_2^2$ as a function of the eigenvalues of $S$.
+Let $\lambdavec(S) = (\lambda_1(S),\ldots,\lambda_n(S)) \in \Reals^n$ denote these eigenvalues, and let $v_k(S)$ denote the corresponding unit-norm eigenvectors, so that $S = \sum_{k = 1}^{n} \lambda_k(S) \cdot v_k(S) v_k(S)^T$.
+Denote $Z_k = v_k(S)^T \varepsilon$, and observe that ${\bf Z} = (Z_1,\ldots,Z_n) \sim N(0,\Id)$.
 
@@
 	\label{lem:linear_smoother_fixed_graph_estimation}
-	Let $\check{f} = SY$ for a square, symmetric matrix, $S \in \Reals^{n \times n}$ satisfying $\lambda_n(S) \leq 1$. Then
-	\begin{equation*}
-	\Pbb_{f_0}\biggl(\frac{1}{n}\bigl\|\check{f} - f_0\bigr\|_2^2 \geq \frac{10}{n} \bigl\|\lambdavec(S)\bigr\|_2^2 + \frac{2}{n}\bigl\|(S - I)f_0\bigr\|_2^2\biggr) \leq 1 - \exp\Bigl(-\bigl\|\lambdavec(S)\bigr\|_2^2\Bigr)
+	Let $\wc{f} = SY$ for a square, symmetric matrix, $S \in \Reals^{n \times n}$. Then
+	\begin{equation*}
+	\Pbb_{f_0}\biggl(\frac{1}{n}\bigl\|\wc{f} - f_0\bigr\|_2^2 \geq \frac{10}{n} \bigl\|\lambdavec(S)\bigr\|_2^2 + \frac{2}{n}\bigl\|(S - I)f_0\bigr\|_2^2\biggr) \leq 1 - \exp\Bigl(-\bigl\|\lambdavec(S)\bigr\|_2^2\Bigr)
 	\end{equation*}
 \end{lemma}
-
-On the testing side, the functional $\norm{\check{f}}_2^2 = Y^T S^2 Y$ is a $U$-statistic of order $2$.
-The statistic $T_{\LS}(G)$ can be written in this form, and Lemma~\ref{lem:linear_smoother_fixed_graph_testing} is therefore useful in analyzing it.
+Here we have written $\Pbb_{f_0}(\cdot)$ for the probability law under the regression ``function'' $f_0 \in \Reals^n$.
+
+In Lemma~\ref{lem:linear_smoother_fixed_graph_testing}, we upper bound the error of a test involving the statistic $\|\wc{f}\|_2^2 = {\bf Y}^T S^2 {\bf Y}$.
 \begin{lemma}
 	\label{lem:linear_smoother_fixed_graph_testing}
-	Let $T = Y^T S^2 Y$ for a square, symmetric matrix $S \in \Reals^{n \times n}$. Define the threshold $t_b$ to be 
-	\begin{equation}
-	t_{b} := \norm{\lambdavec(S)}_2^2 + 2b \norm{\lambdavec(S)}_4^2
-	\end{equation}
-	Suppose the eigenvalues $0 \leq \lambda_{1}(S) \leq \lambda_{n}(S) \leq 1$. Then,
+	Let $\wc{T} = Y^T S^2 Y$ for a square, symmetric matrix $S \in \Reals^{n \times n}$. Define the threshold $\wc{t}_b$ to be 
+	\begin{equation}
+	\wc{t}_{b} := \norm{\lambdavec(S)}_2^2 + 2b \norm{\lambdavec(S)}_4^2
+	\end{equation}
+	It holds that:
 	\begin{itemize}
@@
 		\label{eqn:linear_smoother_fixed_graph_testing_typeI}
-		\Pbb_0\bigl(T > t_b\bigr) \leq \frac{1}{b^2}
-		\end{equation}
-		\item \textbf{Type II error.} Suppose further that
+		\Pbb_0\bigl(\wc{T} > \wc{t}_b\bigr) \leq \frac{1}{b^2}
+		\end{equation}
+		\item \textbf{Type II error.} Under the further assumption
 		\begin{equation}
 		\label{eqn:linear_smoother_fixed_graph_testing_critical_radius}
-		f_0^T S^2 f_0 \geq 4b \norm{\lambdavec(S)}_4^2.
-		\end{equation}
-		Then
+		f_0^T S^2 f_0 \geq 4b \norm{\lambdavec(S)}_4^2,
+		\end{equation}
+		then
 		\begin{equation}
 		\label{eqn:linear_smoother_fixed_graph_testing_typeII}
-		\Pbb_{f_0}\bigl(T \leq t_b\bigr) \leq \frac{4}{b^2} + \frac{8}{b \norm{\lambdavec(S)}_4^{2}} 
+		\Pbb_{f_0}\bigl(\wc{T} \leq \wc{t}_b\bigr) \leq \frac{4}{b^2} + \frac{8}{b \norm{\lambdavec(S)}_4^{2}}.
 		\end{equation}
@@
 \paragraph{Proof of Lemma~\ref{lem:linear_smoother_fixed_graph_estimation}.}
-It holds that
-\begin{align*}
-\frac{1}{n}\bigl\|\check{f} - f_0\bigr\|_2^2 & \leq \frac{2}{n}\Bigl(\bigl\|\check{f} - \Ebb_{f_0}[\check{f}]\bigr\|_2^2 + \bigl\|\Ebb_{f_0}[\check{f}] - f_0\bigr\|_2^2\Bigr) \\ 
+The expectation $\Ebb_{f_0}[\wc{f}] = Sf_0$, and by the triangle inequality,
+\begin{align*}
+\frac{1}{n}\bigl\|\wc{f} - f_0\bigr\|_2^2 & \leq \frac{2}{n}\Bigl(\bigl\|\wc{f} - \Ebb_{f_0}[\wc{f}]\bigr\|_2^2 + \bigl\|\Ebb_{f_0}[\wc{f}] - f_0\bigr\|_2^2\Bigr) \\ 
 & = \frac{2}{n}\Bigl(\bigl\|S\varepsilon\bigr\|_2^2 + \bigl\|(S - I)f_0\bigr\|_2^2\Bigr)
@@
 
-\textit{Mean.} Writing $Y = f_0 + \varepsilon$, we make use of the eigendecomposition of $S$ to obtain
+\textit{Mean.} We make use of the eigendecomposition $S = \sum_{k = 1}^{n} \lambda_k(S) \cdot v_k(S) v_k(S)^T$ to obtain
 \begin{equation}
@@
 \begin{aligned}
-T & = f_0^T S^2 f_0 + 2 f_0^T S^2 \varepsilon + \varepsilon^T S^2 \varepsilon \\
+\wc{T} & = f_0^T S^2 f_0 + 2 f_0^T S^2 \varepsilon + \varepsilon^T S^2 \varepsilon \\
 & = f_0^T S^2 f_0 + 2 f_0^T S^2 \varepsilon + \sum_{k = 1}^{n}  \bigl(\lambda_k(S)\bigr)^2 (\varepsilon^T v_k(S))^2 \\
@@
 \label{pf:linear_smoother_fixed_graph_testing_mean}
-\Ebb_{f_0}[T] = f_0^T S^2 f_0 + \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^2.
-\end{equation}
-
-\textit{Variance.} Starting from~\eqref{pf:linear_smoother_fixed_graph_testing1} and recalling the basic fact $\Var(Z_k^2) = 2$, we derive
+\Ebb_{f_0}\bigl[\wc{T}\bigr] = f_0^T S^2 f_0 + \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^2.
+\end{equation}
+
+\textit{Variance.} Starting from~\eqref{pf:linear_smoother_fixed_graph_testing1} and recalling $\Var(Z_k^2) = 2$, we derive
 \begin{equation}
 \label{pf:linear_smoother_fixed_graph_testing_var}
-\Var_{f_0}[T] \leq 8 f_0^T S^4 f_0 + 4 \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^4 \leq 8 f_0^T S^2 f_0 + 4 \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^2
-\end{equation}
-where the second inequality follows since by assumption $\lambda_{n}(S) \leq 1$.
-
-\textit{Bounding Type I and Type II error.} The bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeI} follows directly from Chebyshev's inequality, along with our above calculations on the mean and variance of $T$.
+\Var_{f_0}\bigl[\wc{T}\bigr] \leq 8 f_0^T S^4 f_0 + 4 \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^4.
+\end{equation}
+
+\textit{Bounding Type I and Type II error.} The upper bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeI} follows immediately from~\eqref{pf:linear_smoother_fixed_graph_testing_mean}, ~\eqref{pf:linear_smoother_fixed_graph_testing_var}, and Chebyshev's inequality.
 
@@
 \begin{aligned}
-\Pbb_{f_0}\bigl(T \leq t_b\bigr) & = \Pbb_{f_0}\bigl(T - \Ebb_{f_0}[T] \leq t_b - \Ebb_{f_0}[T]\bigr) \\
-& \overset{(i)}{\leq} \Pbb_{f_0}\bigl(\abs{T - \Ebb_{f_0}[T]} \geq \abs{t_b - \Ebb_{f_0}[T]}\bigr) \\ 
-& \overset{(ii)}{\leq} 4 \frac{\Var_{f_0}[T]}{(f_0^T S^2 f_0)^2} \\
+\Pbb_{f_0}\Bigl(\wc{T} \leq \wc{t}_b\Bigr) & = \Pbb_{f_0}\Bigl(\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr] \leq \wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr) \\
+& \overset{(i)}{\leq} \Pbb_{f_0}\Bigl(\Bigl|\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr| \geq \Bigl|\wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr|\Bigr) \leq \frac{Var_{f_0}\bigl[\wc{T}\bigr]}{\bigl(\wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\bigr)^2} \\
+& \overset{(ii)}{\leq} 4 \frac{\Var_{f_0}\bigl[\wc{T}\bigr]}{(f_0^T S^2 f_0)^2} \\
 & \overset{(iii)}{\leq} \frac{32}{f_0^T S^2 f_0} + \frac{4}{b^2} \\
@@
 \end{equation*}
-In the previous expression, $(i)$ and $(ii)$ follow since assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and equation~\eqref{pf:linear_smoother_fixed_graph_testing_mean} together imply $\Ebb_{f_0}(T) - \frac{1}{2}f_0^T S^2f_0 \geq t_b$, $(iii)$ follows from assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and the inequality~\eqref{pf:linear_smoother_fixed_graph_testing_var}, and $(iv)$ follows assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}.
+In the previous expression, $(i)$ and $(ii)$ follow since assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and equation~\eqref{pf:linear_smoother_fixed_graph_testing_mean} together imply
+\begin{equation*}
+\wc{t}_b - \Ebb_{f_0}[\wc{T}] = f_0^T S^2 f_0 - 2b\|{\bf \lambda}(S)\|_4^2 \geq \frac{1}{2}f_0^T S^2 f_0
+\end{equation*}
+$(iii)$ follows from assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and the inequality~\eqref{pf:linear_smoother_fixed_graph_testing_var}, and $(iv)$ follows from~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}.
 
 \subsection{Analysis of Laplacian Smoothing}
-In Lemma~\ref{lem:ls_fixed_graph_estimation}, we upper bound the mean squared error of $\wh{f}_{\LS}(G)$.
+Upper bounds on the mean squared error of $\wh{f}$, and Type I and Type II error of $\wh{T}$, follow from setting $S = (\rho L + I)^{-1}$ in Lemmas~\ref{lem:linear_smoother_fixed_graph_estimation} and~\ref{lem:linear_smoother_fixed_graph_testing}.
+We give these results in Lemma~\ref{lem:ls_fixed_graph_estimation} and~\ref{lem:ls_fixed_graph_testing}, and prove them immediately.
+Recall that $\lambda_1,\ldots,\lambda_n$ are the $n$ eigenvalues of $\Lap$ (sorted in ascending order).
 \begin{lemma}
 	\label{lem:ls_fixed_graph_estimation}
-	For any $\rho,s > 0$,
+	For any $\rho > 0$,
 	\begin{equation}
 	\label{eqn:ls_fixed_graph_estimation_prob}
-	\frac{1}{n}\bigl\|\wh{f}_{\LS}(G) - f_0\bigr\|_2^2 \leq \frac{2\rho}{n} \bigl(f_0^T \Lap_{G}^s f_0\bigr) + \frac{10}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k}(G)^s + 1\bigr)^2}
-	\end{equation}
-	with probability at least $1 - \exp\Bigl(-\sum_{k = 1}^{n}\bigl(\rho \lambda_k(G)^s + 1\bigr)^{-2}\Bigr)$.
+	\frac{1}{n}\bigl\|\wh{f} - f_0\bigr\|_2^2 \leq \frac{2\rho}{n} \bigl(f_0^T \Lap f_0\bigr) + \frac{10}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k} + 1\bigr)^2}
+	\end{equation}
+	with probability at least $1 - \exp\Bigl(-\sum_{k = 1}^{n}\bigl(\rho \lambda_k + 1\bigr)^{-2}\Bigr)$.
 \end{lemma}
-
-In Lemma~\ref{lem:ls_fixed_graph_testing}., we upper bound the testing error of a test based on $T_{\LS}(G)$.
+Recall that 
+\begin{equation*}
+\wh{t}_b := \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k + 1\bigr)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k + 1\bigr)^4}}.
+\end{equation*}
 \begin{lemma}
 	\label{lem:ls_fixed_graph_testing}
-	Define the threshold $\wh{t}_b$ to be
-	\begin{equation*}
-	\wh{t}_b := \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k(G)^s + 1\bigr)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k(G)^s + 1\bigr)^4}}
-	\end{equation*}
-	Then each of the following holds for any $\rho,s > 0$ and any $b \geq 1$.
+	For any $\rho > 0$ and any $b \geq 1$, it holds that:
 	\begin{itemize}
@@
 		\label{eqn:ls_fixed_graph_testing_typeI}
-		\Pbb_0\Bigl(T_{\LS}(G) > \wh{t}_b\Bigr) \leq \frac{1}{b^2}
+		\Pbb_0\Bigl(\wh{T} > \wh{t}_b\Bigr) \leq \frac{1}{b^2}
 		\end{equation}
@@
 		\label{eqn:ls_fixed_graph_testing_critical_radius}
-		\frac{1}{n}\norm{f_0}_2^2 \geq \frac{2 \rho}{n} \bigl(f_0^T \Lap_G^s f_0\bigr) + \frac{4b}{n} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G)^s + 1)^4} \Biggr)^{1/2}
+		\frac{1}{n}\norm{f_0}_2^2 \geq \frac{2 \rho}{n} \bigl(f_0^T \Lap f_0\bigr) + \frac{4b}{n} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{1/2}
 		\end{equation}
@@
 		\label{eqn:ls_fixed_graph_testing_typeII}
-		\Pbb_{f_0}\Bigl(T_{\LS}(G) \leq \wh{t}_b\Bigr) \leq \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G)^s + 1)^4} \Biggr)^{-1/2}
+		\Pbb_{f_0}\Bigl(\wh{T}(G) \leq \wh{t}_b\Bigr) \leq \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2}
 		\end{equation}
@@
 \end{lemma}
-
 \paragraph{Proof of Lemma~\ref{lem:ls_fixed_graph_estimation}.}
-Letting $\wh{S} = (\Id + \rho \Lap_G^s)^{-1}$, the estimator $\wh{f}_{\LS}(G) = \wh{S}Y$.
-The matrix $\wh{S} \in \Reals^{n \times n}$ is symmetric, and satisfies $\lambda_k(\wh{S}) \in (0,1)$ for all $k \in [n]$.
-The claim then follows from Lemma~\ref{lem:linear_smoother_fixed_graph_estimation} upon noting first that
-\begin{equation*}
-\bigl\|\lambdavec(\wh{S})\bigr\|_2^2 = \sum_{k = 1}^{n} \frac{1}{\bigl(1 + \rho \lambda_k(G)^s\bigr)^2}
-\end{equation*} 
-and second that
+Letting $\wh{S} = (\Id + \rho \Lap)^{-1}$, the estimator $\wh{f} = \wh{S}Y$, and
+\begin{equation*}
+\bigl\|\lambdavec(\wh{S})\bigr\|_2^2 = \sum_{k = 1}^{n} \frac{1}{\bigl(1 + \rho \lambda_k\bigr)^2}.
+\end{equation*} 
+We deduce the following upper bound on the bias term,
 \begin{equation*}
 \begin{aligned}
-\bigl\|(\wh{S} - I)f_0\bigr\|_2^2 & = f_0^T \Lap_G^{s/2} \Lap_G^{-s/2}\bigl(\wh{S} - \Id\bigr) \Lap_G^{-s/2} \Lap_G^{s/2} f_0 \\
-& \leq f_0^T \Lap_G^s f_0 \cdot \lambda_n\Bigl(\Lap_G^{-s/2}\bigl(\wh{S} - \Id\bigr)\Lap_G^{-s/2}\Bigr) \\
-& = f_0^T \Lap_G^s f_0 \cdot \max_{k \in [n]} \biggl\{\frac{1}{\lambda_k(G)^s} \Bigl(1 - \frac{1}{\rho\lambda_k(G)^s + 1}\Bigr) \biggr\} \\
-& = f_0^T \Lap_G^s f_0 \cdot \rho.
+\bigl\|(\wh{S} - I)f_0\bigr\|_2^2 & = f_0^T \Lap^{1/2} \Lap^{-1/2}\bigl(\wh{S} - \Id\bigr) \Lap^{-1/2} \Lap^{1/2} f_0 \\
+& \leq f_0^T \Lap f_0 \cdot \lambda_n\Bigl(\Lap^{-1/2}\bigl(\wh{S} - \Id\bigr)\Lap^{-1/2}\Bigr) \\
+& = f_0^T \Lap f_0 \cdot \max_{k \in [n]} \biggl\{\frac{1}{\lambda_k} \Bigl(1 - \frac{1}{\rho\lambda_k + 1}\Bigr) \biggr\} \\
+& = f_0^T \Lap f_0 \cdot \rho.
 \end{aligned}
 \end{equation*} 
-where we write $\Lap_G^{-1}$ for the pseudoinverse of $\Lap_G$.
+In the above, we have written $\Lap^{-1/2}$ for the square root of the pseudoinverse of $\Lap$, and the maximum is over all indices $k$ such that $\lambda_k > 0$.
 
 \paragraph{Proof of Lemma~\ref{lem:ls_fixed_graph_testing}.}
-Let $\wh{S} := (\rho \Lap^s + \Id)^{-1}$.
-The matrix $\wh{S} \in \Reals^{n \times n}$ is symmetric and positive semidefinite, and our test statistic $T_{\LS}(G) = \frac{1}{n}Y^T \wh{S}^2 Y$.
-The desired result thus follows from Lemma~\ref{lem:linear_smoother_fixed_graph_testing}.
-To see that the conditions of this Lemma are satisfied, we first note that since
-\begin{equation*}
-\lambda_k(\wh{S}) = \frac{1}{(\rho\lambda_k(G)^s + 1)}
-\end{equation*}
-and $\rho, \lambda_k(G) > 0$, it is evident that $\lambda_{n}(\wh{S}) \leq 1$.  Then, by assumption~\eqref{eqn:ls_fixed_graph_testing_critical_radius}
-\begin{equation*}
-f_0^{T} \wh{S}^2 f_0 = \norm{f_0}_2^2 - f_0^T(I - \wh{S}^2)f_0 \geq 2 \rho (f_0^T \Lap^s f_0) + f_0^T(I - \wh{S}^2)f_0 + 4b \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G)^s + 1)^4} \Biggr)^{-1/2},
-\end{equation*}
-and along with the following calculations,
+Let $\wh{S} := (\rho \Lap + \Id)^{-1}$, and note that $\wh{T} = \frac{1}{n}{\bf Y}^T \wh{S}^2 {\bf Y}$.
+The bound on Type I error~\eqref{eqn:ls_fixed_graph_testing_typeI} follows immediately from \eqref{eqn:linear_smoother_fixed_graph_testing_typeI}.
+To establish the bound on Type II error, we must lower bound $f_0^T \wh{S}^2 f_0$. We first note that by assumption~\eqref{eqn:ls_fixed_graph_testing_critical_radius},
+\begin{align*}
+f_0^{T} \wh{S}^2 f_0 & = \bigl\|f_0\bigr\|_2^2 - f_0^T(I - \wh{S}^2)f_0 \\
+& \geq 2 \rho \bigl(f_0^T \Lap f_0\bigr) - f_0^T\bigl(I - \wh{S}^2\bigr)f_0 + 4b \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2}.
+\end{align*}
+Upper bounding $f_0^T\bigl(I - \wh{S}^2\bigr)f_0$ as follows:
 \begin{equation*}
 \begin{aligned}
-f_0^T \Bigl(\Id - \wh{S}^2\Bigr) f_0  & = f_0^T L^{s/2} L^{-s/2}\Bigl(\Id - \wh{S}^2\Bigr) L^{-s/2} L^{s/2} f_0 \\ 
-& \leq f_0^T L^{s} f_0 \cdot  \lambda_{\max}\biggl(L^{-s/2}\Bigl(\Id - \wh{S}^2\Bigr) L^{-s/2}\biggr) \\ 
-& \overset{(i)}{=}  f_0^T L^{s} f_0 \cdot \max_{k} \biggl\{ \frac{1}{\lambda_k(G)^s} \Bigl(1 - \frac{1}{(\rho \lambda_k(G)^s + 1)^2}\Bigr) \biggr\} \\
-& \overset{(ii)}{\leq} f_0^T L^{s} f_0 \cdot 2\rho,
+f_0^T \Bigl(\Id - \wh{S}^2\Bigr) f_0  & = f_0^T \Lap^{1/2} \Lap^{-1/2}\Bigl(\Id - \wh{S}^2\Bigr) \Lap^{-1/2} \Lap^{1/2} f_0 \\ 
+& \leq f_0^T \Lap f_0 \cdot  \lambda_{n}\biggl(\Lap^{-1/2}\Bigl(\Id - \wh{S}^2\Bigr) \Lap^{-1/2}\biggr) \\ 
+& \overset{(i)}{=}  f_0^T \Lap f_0 \cdot \max_{k} \biggl\{ \frac{1}{\lambda_k} \Bigl(1 - \frac{1}{(\rho \lambda_k + 1)^2}\Bigr) \biggr\} \\
+& \overset{(ii)}{\leq} f_0^T \Lap f_0 \cdot 2\rho,
 \end{aligned}
 \end{equation*}
-we have that
-\begin{equation*}
-f_0^{T} \wh{S}^2 f_0 \geq 2b \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G)^s + 1)^4} \biggr)^{-1/2}.
-\end{equation*} 
-In other words condition~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} in Lemma~\ref{lem:linear_smoother_fixed_graph_testing} is met, and applying that Lemma completes the proof.
-
-(In the previous derivation: in $(i)$ the maximum is over all indices $k$ such that the eigenvalue $\lambda_k(G)$ is strictly positive; and $(ii)$ follows from the basic algebraic identity $1 - 1/(1 + \rho x)^2 \leq 2 \rho x$ for any $x, \rho > 0$.
+---where in the above the maximum is over all indices $k$ such that $\lambda_k > 0$, and the last inequality follows from the basic algebraic identity $1 - 1/(1 + s x)^2 \leq 2 s x$ for any $x, s > 0$---we deduce that
+\begin{equation*}
+f_0^{T} \wh{S}^2 f_0 \geq 2b \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{-1/2}.
+\end{equation*} 
+The upper bound on Type II error~\eqref{eqn:ls_fixed_graph_testing_typeII} then follows from Lemma~\ref{lem:linear_smoother_fixed_graph_testing}.
 
@@
 \label{sec:graph_sobolev_seminorm}
-In this section, we prove Lemma~\ref{lem:graph_sobolev_seminorm}. To do so, we will show that for any $f \in H^1(\Xset)$,
+In this section, we prove Lemma~\ref{lem:graph_sobolev_seminorm}, which states an upper bound on $f^T \Lap_{n,r} f$ holds when $f$ is bounded in Sobolev norm.
+We also establish stronger bounds in the case when $f$ has a bounded Lipschitz constant; this latter result justifies the remark one of our remarks after Theorem~\ref{thm:laplacian_smoothing_estimation1}.
+
+To prove Lemma~\ref{lem:graph_sobolev_seminorm}, we will show that for any $f \in H^1(\Xset)$,
 \begin{equation*}
@@
 
-For the remainder of this proof, we will assume that $f \in C^{\infty}(\Xset)$, which we may do without loss of generality because $C^{\infty}(\Xset)$ is dense in $H^1(\Xset)$ and the expectation on the right hand side of~\eqref{pf:first_order_graph_sobolev_seminorm_1} is continuous in $\Leb^2(\Xset)$.
+For the remainder of this proof, we will assume that $f \in C^{\infty}(\Xset)$.
+We may do so without loss of generality: $C^{\infty}(\Xset)$ is dense in $H^1(\Xset)$ and the expectation on the right hand side of~\eqref{pf:first_order_graph_sobolev_seminorm_1} is continuous in $H^1(\Xset)$.
 Obviously
@@
 
+\subsection{Stronger bounds under Lipschitz assumption}
+Suppose $f$ satisfies $|f(x') - f(x)| \leq M \|x - x\|$ for all $x,x' \in \mc{X}$.
+Then we can strengthen the high probability bound in Lemma~\ref{lem:graph_sobolev_seminorm} from $1 - \delta$ to $1 - \delta^2/n$, at the cost of only a constant factor in the upper bound on $f^T \Lap_{n,r} f$.
+\begin{proposition}
+	\label{prop:graph_sobolev_seminorm_lipschitz}
+	Let $r \geq C_0(\log n/n)^{1/d}$. For any $f$ such that $|f(x') - f(x)| \leq M \|x - x\|$ for all $x,x' \in \mc{X}$, with probability at least $1 - C\delta^2/n$ it holds that
+	\begin{equation*}
+	f^T \Lap_{n,r} f \leq \biggl(\frac{1}{\delta} + (1 + L_p)^2 \sigma_K\biggr) n^2 r^{d + 2} M^2
+	\end{equation*}
+\end{proposition}
+\paragraph{Proof of Proposition~\ref{prop:graph_sobolev_seminorm_lipschitz}.}
+We will prove Proposition~\ref{prop:graph_sobolev_seminorm_lipschitz} using Chebyshev's inequality, so the key step is to upper bound the variance of $f^{T}\Lap_{n,r}f$.
+Putting $\varDelta_{ij} := K(\|X_i - X_j\|/r) \cdot (f(X_i) - f(X_j))^2$, we can write the variance of $f^T \Lap_{n,r} f$ as a sum of covariances,
+\begin{equation*}
+\Var\bigl[f^T \Lap_{n,r} f\bigr] = \frac{1}{4} \sum_{i, j = 1}^{n} \sum_{\ell,m = 1}^{n} \Cov\bigl[\varDelta_{ij},\varDelta_{\ell m}\bigr].
+\end{equation*}
+Clearly $\Cov\bigl[\varDelta_{ij},\varDelta_{\ell m}\bigr]$ depends on the cardinality of $I := \{i,j,k,\ell\}$; we divide into cases, and upper bound the covariance in each case.
+\begin{enumerate}
+	\item[$\bigl|I\bigr| = 4$.] In this case $\varDelta_{ij}$ and $\varDelta_{\ell m}$ are independent, and $\Cov\bigl[\varDelta_{ij},\varDelta_{\ell m}\bigr] = 0$.
+	\item[$\bigl|I\bigr| = 3$.] Taking $i = \ell$ without loss of generality, we have
+	\begin{align*}
+	\Cov\bigl[\varDelta_{ij},\varDelta_{i m}\bigr] & \leq \Ebb\bigl[\varDelta_{ij} \varDelta_{i m} \bigr] \\
+	& = \int_{\Xset} \int_{\Xset} \int_{\Xset} \bigl(f(z) - f(x)\bigr)^2 \bigl(f(x') - f(x)\bigr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) K\biggl(\frac{\|z - x\|}{r}\biggr) p(z) p(x') p(x) \,dz \,dx' \,dx \\
+	& \leq p_{\max}^3 M^4 r^4 \int_{\Xset} \int_{\Xset} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr) K\biggl(\frac{\|z - x\|}{r}\biggr) \,dz \,dx' \,dx \\
+	& \leq p_{\max}^3 M^4 r^{4 + 2d}.
+	\end{align*}
+	\item[$\bigl|I\bigr| = 2$.] Taking $i = \ell$ and $j = m$ without loss of generality, we have
+	\begin{align*}
+	\Var\bigl[\varDelta_{ij}\bigr] & \leq \Ebb\bigl[\varDelta_{ij}^2\bigr] \\
+	& \leq \int_{\Xset} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^4 \biggl[K\biggl(\frac{\|x' - x\|}{r}\biggr)\biggr]^2 p(x') p(x) \,dx' \,dx \\
+	& \leq p_{\max}^2 M^4 r^4 K(0) \int_{\Xset} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx' \,dx \\
+	& \leq p_{\max}^2 M^4 r^{4 + d} K(0)
+	\end{align*}
+	\item[$\bigl|I\bigr| = 1$.] In this case $\varDelta_{i j} = \varDelta_{\ell m} = 0$.
+\end{enumerate}
+Therefore
+\begin{equation*}
+\Var\bigl[f^T \Lap_{n,r} f\bigr] \leq n^3 p_{\max}^3 M^4 r^{4 + 2d} + n^2 p_{\max}^2 M^4 r^{4 + d} K(0) \leq Cn^3r^{4 + 2d},
+\end{equation*}
+where the latter inequality follows since $nr^d \gg 1$. For any $\delta > 0$, it follows from Chebyshev's inequality that
+\begin{equation*}
+\Pbb\biggl(\Bigl|f^T \Lap_{n,r} f - \Ebb\bigl[f^T \Lap_{n,r} f\bigr]\Bigr| \geq \frac{M^2}{\delta}n^2 r^{d + 2}\biggr) \leq C\frac{\delta^2}{n}
+\end{equation*}
+and since we have already upper bounded $\Ebb\bigl[f^T \Lap_{n,r} f\bigr] \leq (1 + L_p)^2 \sigma_K M^2 n^2 r^{d + 2}$, the proposition follows.
+
+Note that the bound on $\Var[\varDelta_{i j}]$ follows as long as we can control $\|\nabla f\|_{\Leb^4(\Xset)}$; this implies the Lipschitz assumption---which gives us control of $\|\nabla f\|_{\Leb^{\infty}(\Xset)}$---can be weakened.
+However, the Sobolev assumption---which gives us control only over $\|\nabla f\|_{\Leb^2(\Xset)}$---will not do the job.
 \section{Bounds on neighborhood graph eigenvalues}
 \label{sec:graph_eigenvalues}
-In this section, we prove our claimed results regarding the eigenvalues $\lambda_k(G_{n,r})$.
-We begin with Corollary~\ref{cor:neighborhood_eigenvalue}, since its proof is straightforward, and then turn to the proof of Lemma~\ref{lem:neighborhood_eigenvalue}, which is more involved.
-
-\subsection{Proof of Corollary~\ref{cor:neighborhood_eigenvalue}}
+In this section, we prove Lemma~\ref{lem:neighborhood_eigenvalue}, following the lead of~\citep{burago2014,trillos2019,calder2019}, who establish similar results with respect to a manifold without boundary.
+The bulk of our technical effort will stem from proving Theorem~\ref{thm:neighborhood_eigenvalue}, which gives estimates on eigenvalues of the graph Laplacian in terms of eigenvalues of the weighted Laplace-Beltrami operator $\Delta_P$.
+We recall $\Delta_P$ is defined as
+\begin{equation*}
+\Delta_Pf(x) := -\frac{1}{p} \dive\bigl(p^2\nabla f(x)\bigr)
+\end{equation*}
+To avoid confusion, in this section we write $\lambda_k(G_{n,r})$ for the $k$th smallest eigenvalue of the graph Laplacian matrix $\Lap_{n,r}$ and $\lambda_k(\Delta_P)$ for the $k$th smallest eigenvalue of $\Delta_P$ \footnote{Under the assumptions}.
+Some other notation: throughout this section, we will write $A, A_1,\ldots$ and $a,a_1,\ldots$ for constants which do not depend on $n$, and we let $\theta$ and $\wt{\delta}$ denote (small) positive numbers, which always satisfy the following restrictions.
+\begin{enumerate}[label=(S\arabic*)]
+	\item 
+	\begin{equation*}
+	\theta + \wt{\delta} \leq \frac{1}{4A_1}~~\textrm{and}~~C_0\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq \wt{\delta} \leq r
+	\end{equation*}
+\end{enumerate}
+
+\begin{theorem}
+	\label{thm:neighborhood_eigenvalue}
+	For any $\ell \in \mathbb{N}$ and $\theta$ and $\wt{\delta}$ such that
+	\begin{equation}
+	\label{eqn:neighborhood_eigenvalue_1}
+	1 - A\Biggl(r \sqrt{\lambda_{\ell}(\Delta_P)} + \theta + \wt{\delta}\Biggr)\geq \frac{1}{2}
+	\end{equation}
+	with probability at least $1 - A_1n\exp(-a_1n\theta^2\wt{\delta}^{m})$, it holds that
+	\begin{equation}
+	\label{eqn:eigenvalue_bound}
+	a \lambda_k(G_{n,r}) \leq nr^{d+2} \lambda_k(\Delta_P) \leq A \lambda_k(G_{n,r}),~~\textrm{for all $1 \leq k \leq \ell$}
+	\end{equation}
+\end{theorem}
+We will return to proving Theorem~\ref{thm:neighborhood_eigenvalue}, but first we show that under our regularity conditions on $p$ and $\Xset$, it implies Lemma~\ref{lem:neighborhood_eigenvalue}.
+The link between the two is Weyl's Law.
+\begin{proposition}[Weyl's Law]
+	\label{prop:weyl}
+	Suppose the density $p$ and the domain $\Xset$ satisfy~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density}. Then there exist constants $a_2$ and $A_2$ such that
+	\begin{equation}
+	\label{eqn:weyls_law}
+	a_2k^{2/d} \leq \lambda_k(\Delta_P) \leq A_2k^{2/d}~~\textrm{for all $k \in \mathbb{N}, k > 1$}.
+	\end{equation}
+\end{proposition}
+See Lemma 28 of~\citep{dunlop2020} for a proof that~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density} imply Weyl's Law.
+\paragraph{Proof of Lemma~\ref{lem:neighborhood_eigenvalue}.}
 Put
 \begin{equation*}
-\ell_{\star} = \floor{\frac{\bigl(1/2 - A_1(\theta + \wt{\delta})\bigr)^d}{C_2^d A_1^d r^d}};
-\end{equation*}
-and note that for $\ell_{\star}$ the condition~\eqref{eqn:neighborhood_eigenvalue_1} is met, i.e.
-\begin{equation*}
-1 - A_1\Biggl(r \sqrt{\lambda_{\ell}(\Delta_P)} + (\theta + \wt{\delta})\Biggr)  \geq \frac{1}{2}.
-\end{equation*}
-We may therefore apply the conclusions of Lemma~\ref{lem:neighborhood_eigenvalue}, which along with~\eqref{eqn:weyls_law} imply that
-\begin{equation*}
-\frac{c_2}{A_1} nr^{d + 2} k^{2/d} \leq \lambda_k(G_{n,r}) \leq \frac{C_2}{a_1} nr^{d + 2} k^{2/d}~~\textrm{for all $2 \leq k \leq \ell_{\star}$}
-\end{equation*}
-On the other hand, if $k > \ell^{\star}$,...
-
-\textcolor{red}{(TODO)}: Finish off the proof.
-
-\subsection{Proof of Lemma~\ref{lem:neighborhood_eigenvalue}}
-
-In this section we prove Lemma~\ref{lem:neighborhood_eigenvalue}, closely following the lead of~\citep{burago2014,trillos2019,calder2019}.
-As in these works, we relate $\lambda_k(\Delta_P)$ and $\lambda_k(G_{n,r})$ by means of the Sobolev semi-norms (Dirichlet energies)
-\begin{equation*}
-b_r(u) := \frac{1}{n^2 r^{d+ 2}}u^T \Lap_{n,r} u~~\textrm{for any $u \in \Leb^2(P_n)$}
+\ell_{\star} = \floor{\biggl(\frac{\bigl(1/(2A_1) - (\theta + \wt{\delta})\bigr)}{rA_2^{1/2}}\biggr)^d},
+\end{equation*}
+and note that the condition $r < c_0$ guarantees $\ell_{\star} \geq 2$ when $c_0$ is chosen to be a sufficiently small constant. By Proposition~\ref{prop:weyl}, we have that
+\begin{align*}
+\sqrt{\lambda_{\ell_{\star}}(\Delta)} \leq \frac{1}{r}\biggl(\frac{1}{2A_1} - (\theta + \wt{\delta})\biggr) 
+\end{align*}
+so that condition~\eqref{eqn:neighborhood_eigenvalue_1} is satisfied.
+We will therefore assume the conclusions of Theorem~\ref{thm:neighborhood_eigenvalue}.
+Along with~\eqref{eqn:weyls_law}, this implies the following bounds on the graph Laplacian eigenvalues:
+\begin{equation*}
+\frac{a_2}{A_1} nr^{d + 2} k^{2/d} \leq \lambda_k(G_{n,r}) \leq \frac{A_2}{a_1} nr^{d + 2} k^{2/d}~~\textrm{for all $2 \leq k \leq \ell_{\star}$}.
+\end{equation*}
+It remains to bound $\lambda_k(G_{n,r})$ for those indices $k$ which are greater than $\ell_{\star}$.
+On the one hand, since the eigenvalues are sorted in ascending order, we can use the lower bound on $\lambda_{\ell_{\star}}(G_{n,r})$ that we have just derived:
+\begin{align*}
+\lambda_k(G_{n,r}) & \geq \lambda_{\ell_{\star}}(G_{n,r}) \\
+& \geq \frac{a_2}{A_1}nr^{d + 2}\ell_{\star}^{2/d} \\
+& = \frac{a_2}{A_1}nr^{d + 2} \floor{\biggl(\frac{\bigl(1/(2A_1) - (\theta + \wt{\delta})\bigr)}{rA_2^{1/2}}\biggr)^d}^2 \\
+& \geq \frac{a_2}{32A_1^3 A_2} nr^{d}
+\end{align*}
+with the last line following since $\theta + \wt{\delta} \leq \frac{1}{4A_1}$.
+On the other hand, for any graph $G$ the maximum eigenvalue of the Laplacian is upper bounded by twice the maximum degree \citep{chung97}.
+Writing $D_{\max}(G_{n,r})$ for the maximum degree of $G_{n,r}$, we show in Lemma~\ref{lem:max_degree} that
+\begin{equation*}
+\lambda_k(G_{n,r}) \leq 2D_{\max}(G_{n,r}) \leq 4p_{\max}nr^d 
+\end{equation*}
+with probability at least $1 - 2n\exp\Bigl(-nr^dp_{\min}/(3K(0)^2)\Bigr)$. Thus we have established each of the inequalities in Lemma~\ref{lem:neighborhood_eigenvalue}.
+
+\subsection{Proof of Theorem~\ref{thm:neighborhood_eigenvalue}}
+
+In this section we prove Theorem~\ref{thm:neighborhood_eigenvalue}, following closely the approach of \citep{burago2014,trillos2019,calder2019}.
+As in these works, we relate $\lambda_k(\Delta_P)$ and $\lambda_k(G_{n,r})$ by means of the Dirichlet energies
+\begin{equation*}
+b_r(u) := \frac{1}{n^2 r^{d+ 2}}u^T \Lap_{n,r} u 
 \end{equation*}
@@
 \end{equation*}
-Let us pause briefly to motivate the relevance of $b_r(u)$ and $D_2(f)$.
-Suppose one could upper bound $b_r(u)$ by $D_2\bigl(\mc{I}(u)\bigr)$ for an appropriate choice of interpolating map $\mc{I}: \Leb^2(P_n) \to \Leb^2(P)$, and vice versa upper bound $D_2(f)$ by $b_r(\mc{P}(f))$ for an appropriate choice of discretization map $\mc{P}: \Leb^2(P) \to \Leb^2(P_n)$.
-Suppose further that $\mc{I}$ and $\mc{P}$ were near-isometries, meaning $\|\mc{I}(u)\|_{\Leb^2(P)} \approx \|u\|_{\Leb^2(P_n)}$ and $\|\mc{P}(f)\|_{\Leb^2(P_n)} \approx \|f\|_{\Leb^2(P)}$.
-Then using the variational characterization of eigenvalues $\lambda_k(\Delta_P)$ and $\lambda_k(G_{n,r})$---i.e.
-the Courant-Fischer Theorem---one could obtain estimates on the error of $|\lambda_k(\Delta_P) - \lambda_k(G_{n,r})|$, and thereby establish Lemma~\ref{lem:neighborhood_eigenvalue}.
-
-We will momentarily define the maps $\wt{\mc{I}}$ and $\wt{\mc{P}}$, and subsequently prove they satisfy the outlined desiderata.
-These maps will be defined with respect to a particular probability measure $\wt{P}_n$ that with high probability is close in transportation distance to both $P_n$ and $P$---this probabilistic estimate is the workhorse that allows us to relate the random energy $b_r$ to the non-random $D_2$.
-We now review the notion of transportation distance, as well formalize the properties that $\wt{P}_n$ will satisfy.
+We recall that for a function $u: \{X_1,\ldots,X_n\} \to \Reals$, the empirical norm is defined as $\|u\|_n^2 := \frac{1}{n} \sum_{i = 1}^{n} (u(X_i))^2$, and we let the class $\Leb^2(P_n)$ consist of those $u \in \Reals^n$ for which $\|u\|_{n} < \infty$.
+For a function $f: \Xset \to \Reals$, the continuum limit of $\|f\|_n^2$ is simply
+\begin{equation*}
+\|f\|_{P}^2 := \int_{\Xset} \bigl|f(x)\bigr|^2 p(x) \,dx,
+\end{equation*} 
+and we let $\Leb^2(P)$ be the set of functions for which $\|f\|_P < \infty$.
+
+Let us pause briefly to motivate the relevance of $b_r(u)$ and $D_2(f)$. Suppose one could show the following two results: 
+\begin{enumerate}[(1)]
+	\item an upper bound of $b_r(u)$ by $D_2\bigl(\mc{I}(u)\bigr)$ for an appropriate choice of interpolating map $\mc{I}: \Leb^2(P_n) \to \Leb^2(\mc{X})$, and vice versa upper bound $D_2(f)$ by $b_r(\mc{P}(f))$ for an appropriate choice of discretization map $\mc{P}: \Leb^2(\mc{X}) \to \Leb^2(P_n)$,
+	\item that $\mc{I}$ and $\mc{P}$ were near-isometries, meaning $\|\mc{I}(u)\|_{P} \approx \|u\|_{n}$ and $\|\mc{P}(f)\|_{P} \approx \|f\|_{n}$.
+\end{enumerate}
+Then, by using the variational characterization of eigenvalues $\lambda_k(\Delta_P)$ and $\lambda_k(G_{n,r})$---i.e.
+the Courant-Fischer Theorem---one could obtain estimates on the error $\bigl|nr^{d + 2}\lambda_k(\Delta_P) - \lambda_k(G_{n,r})\bigr|$.
+
+We will momentarily define maps $\wt{\mc{I}}$ and $\wt{\mc{P}}$ that satisfy both (1) and (2).
+These maps will be defined with respect to a particular probability measure $\wt{P}_n$, that with high probability is close in transportation distance to both $P_n$ and $P$.
+This estimate on the transportation distance---which we now give--- will be the workhorse that allows us to relate the random energy $b_r$ to the non-random $D_2$.
 
@@
 \end{equation*}
-\cite{calder2019} consider $\Xset$ a smooth manifold without boundary, i.e.
+\cite{calder2019} take $\Xset$ to be a smooth submanifold of $\Rd$ without boundary, i.e.
 they assume $\Xset$ satisfies~\ref{asmp:domain_manifold}.
@@
 In Proposition~\ref{prop:optimal_transport}, we adapt this result to  the setting of full-dimensional manifolds with boundary.
-\begin{proposition}[Restatement of Proposition~2.11 of \textcolor{red}{(Calder2019)}]
+\begin{proposition}
 	\label{prop:optimal_transport}
 	Suppose $p$ satisfies~\ref{asmp:bounded_lipschitz_density} and $\Xset$ satisfies~\ref{asmp:domain}.
-Then with probability at least $1 - A_2 n \exp\bigl\{-a_2 n\theta^2\wt{\delta}^d\bigr\}$, the following statement holds: there exists a probability measure $\wt{P}_n$ with density $\wt{p}_n$ such that:
-	\begin{equation*}
-	d_{\infty}(\wt{P}_n, P_n) \leq A_2 \wt{\delta}
+Then with probability at least $1 - A_1 n \exp\bigl\{-a_1 n\theta^2\wt{\delta}^d\bigr\}$, the following statement holds: there exists a probability measure $\wt{P}_n$ with density $\wt{p}_n$ such that:
+	\begin{equation*}
+	d_{\infty}(\wt{P}_n, P_n) \leq A_1 \wt{\delta}
 	\end{equation*}
@@
 	\begin{equation*}
-	\|\wt{p}_n - p\|_{\infty} \leq A_2\bigl(\wt{\delta} + \theta\bigr)
+	\|\wt{p}_n - p\|_{\infty} \leq A_1\bigl(\wt{\delta} + \theta\bigr)
 	\end{equation*}
@@
 
-The discretization map  $\wt{\mathcal{P}}: \Leb^2(P) \to \Leb^2(P_n)$ is given by
+The discretization map  $\wt{\mathcal{P}}: \Leb^2(\Xset) \to \Leb^2(P_n)$ is given by
 \begin{equation*}
@@
 	\label{prop:dirichlet_energies}
-	With probability at least $1 - A_2n\exp(-a_2n\theta^2\wt{\delta}^{m})$, we have the following.
+	With probability at least $1 - A_1n\exp(-a_1n\theta^2\wt{\delta}^{m})$, we have the following.
 	\begin{enumerate}[(1)]
@@
 	\label{prop:isometry}
-	With probability at least $1 - A_2n\exp(-a_2n\theta^2\wt{\delta}^{m})$, we have the following.
+	With probability at least $1 - A_1n\exp(-a_1n\theta^2\wt{\delta}^{m})$, we have the following.
 	\begin{enumerate}[(1)]
@@
 		\begin{equation*}
-		\Bigl|\|f\|_{\Leb^2(P)}^2 - \|\wt{P}f\|_{\Leb^2(P_n)}^2\Bigr| \leq A_5 \wt{\delta} \|f\|_{\Leb^2(P)} \sqrt{D_2(f)} + \frac{A_1\bigl(\theta + \wt{\delta}\bigr)}{p_{\min}} \|f\|_{\Leb^2(P)}^2
+		\Bigl|\|f\|_{P}^2 - \|\wt{P}f\|_{n}^2\Bigr| \leq A_5 \wt{\delta} \|f\|_{P} \sqrt{D_2(f)} + \frac{A_1\bigl(\theta + \wt{\delta}\bigr)}{p_{\min}} \|f\|_{P}^2
 		\end{equation*}
@@
 		\begin{equation*}
-		\Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(P)}^2 - \|u\|_{\Leb^2(P_n)}^2\Bigr| \leq C r \|u\|_{\Leb^2(P_n)} \sqrt{b_r(u)} + C\bigl(\theta + \wt{\delta}\bigr) \|u\|_{\Leb^2(P_n)}^2
+		\Bigl|\|\wt{\mc{I}}u\|_{P}^2 - \|u\|_{n}^2\Bigr| \leq A_6 r \|u\|_{n} \sqrt{b_r(u)} + A_7\bigl(\theta + \wt{\delta}\bigr) \|u\|_{n}^2
 		\end{equation*}
@@
 
-Given these propositions, the proof of Lemma~\ref{lem:neighborhood_eigenvalue} follows by the Courant-Fischer Theorem.
-
-\paragraph{Proof of Lemma~\ref{lem:neighborhood_eigenvalue}.}
+Given these propositions, the proof of Theorem~\ref{thm:neighborhood_eigenvalue} follows by the Courant-Fischer Theorem, as we now show.
+
+\paragraph{Proof of Theorem~\ref{thm:neighborhood_eigenvalue}.}
 We start with the upper bound, proceeding exactly as in Proposition 4.4 of \citep{burago2014}.
-Let $W$ be the span of $f_1,\ldots,f_{k}$, the first $k$ eigenfunctions of $-\Delta_P$, so that by the Courant-Fischer principle $D_2(f) \leq \lambda_k(\Delta_P) \|f\|_{\Leb^2(P)}$ for every $f \in W$.
+Let $f_1,\ldots,f_{k}$ denote the first $k$ eigenfunctions of $\Delta_P$ and set $W := \mathrm{span}(f_1,\ldots,f_k)$, so that by the Courant-Fischer principle $D_2(f) \leq \lambda_k(\Delta_P) \|f\|_{P}^2$ for every $f \in W$.
 As a result, by Part (1) of Proposition~\ref{prop:isometry} we have that for any $f \in W$
 \begin{equation*}
-\bigl\|\wt{\mc{P}}f\bigr\|_{\Leb^2(P_n)}^2 \geq \biggl(1 - A_5\wt{\delta} \sqrt{\lambda_{k}(\Delta_P)} - \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr)\|f\|_{\Leb^2(P)}^2  \geq \frac{1}{2} \|f\|_{\Leb^2(P)}^2,
+\bigl\|\wt{\mc{P}}f\bigr\|_{n}^2 \geq \biggl(1 - A_5\wt{\delta} \sqrt{\lambda_{k}(\Delta_P)} - \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr)\|f\|_{P}^2  \geq \frac{1}{2} \|f\|_{P}^2,
 \end{equation*}
@@
 \begin{align*}
-\lambda_k(G_{n,r}) & \leq \max_{\substack{u \in \wt{\mc{P}}W \\ u \neq 0} } \frac{b_r(u)}{\|u\|_{\Leb^2(P_n)}^2} \\
-& = \max_{\substack{f \in W \\ f \neq 0} } \frac{b_r(\wt{\mc{P}}f)}{\bigl\|\wt{\mc{P}}f\bigr\|_{\Leb^2(P_n)}^2} \\
-& \leq \frac{\Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + L_p(2r + 4\wt{\delta})\Bigr) \cdot \Bigl(1 + \frac{2L_K}{K(1/2)}\frac{\wt{\delta}}{r}\Bigr)}{1 - A_5\wt{\delta} \sqrt{\lambda_{\ell}(\Delta_P)} - 2\frac{A_1(\theta + \wt{\delta})}{p_{\min}}} \sigma_K \lambda_{\ell}(\Delta_P),
+\frac{\lambda_k(G_{n,r})}{nr^{d + 2}} & \leq \max_{\substack{u \in \wt{\mc{P}}W \\ u \neq 0} } \frac{b_r(u)}{\|u\|_{n}^2} \\
+& = \max_{\substack{f \in W \\ f \neq 0} } \frac{b_r(\wt{\mc{P}}f)}{\bigl\|\wt{\mc{P}}f\bigr\|_{n}^2} \\
+& \leq 2\Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + L_p(2r + 4\wt{\delta})\Bigr) \cdot \Bigl(1 + A_4\frac{\wt{\delta}}{r}\Bigr)\sigma_K \lambda_{k}(\Delta_P),
+\end{align*}
+establishing the lower bound in~\eqref{eqn:eigenvalue_bound} upon proper choice of $a$.
+
+The upper bounds follows from essentially parallel reasoning.
+Recalling that $v_1,\ldots,v_k$ denote the first $k$ eigenvecotrs of $\Lap_{n,r}$, let $U := \mathrm{span}(v_1,\ldots,v_k)$, so that $nr^{d + 2} b_r(u) \leq \lambda_k(G_{n,r}) \|u\|_n^2$.
+By Proposition~\ref{prop:isometry}, Part (2), for every $u \in U$ it holds that
+\begin{align*}
+\bigl\|\wt{\mc{I}}u\bigr\|_{P}^2 & \geq \|u\|_n^2 - A_6 r \|u\|_n \sqrt{b_r(u)} - A_7\bigl(\theta + \wt{\delta}\bigr)\|u\|_n^2 \\
+& \geq \|u\|_n^2 - A_6 r \|u\|_n^2 \sqrt{\frac{\lambda_{k}(G_{n,r})}{nr^{d + 2}}} - A_7\bigl(\theta + \wt{\delta}\bigr)\|u\|_n^2 \\
+& \geq \|u\|_n^2 - A_6 r \|u\|_n^2 \sqrt{\frac{1}{a}\lambda_k(\Delta_P)} - A_7\bigl(\theta + \wt{\delta}\bigr)\|u\|_n^2 \\
+& \geq \frac{1}{2}\|u\|_n^2
+\end{align*}
+where the second to last inequality follows from the lower bound $a \lambda_k(G_{n,r}) \leq nr^{d + 2}\lambda_k(\Delta_P)$ that we just derived, and the last inequality from assumption~\eqref{eqn:neighborhood_eigenvalue_1}.
+
+Therefore $\wt{\mc{I}}$ is injective over $U$, $\wt{\mc{I}}U$ has dimension $k$, and by Proposition~\ref{prop:dirichlet_energies} we conclude that
+\begin{align*}
+\lambda_k(\Delta_P) & \leq \max_{u \in U} \frac{D_2(\wt{\mc{I}}u)}{\|u\|_P^2} \\
+& \leq 80\frac{(1 + L_pr)}{\sigma_K}\biggl(1 + A_1(\theta + \wt{\delta})\biggr) \biggl(1 + A_3\Bigl(\theta + \frac{\wt{\delta}}{r}\Bigr)\biggr)\max_{u \in U} \frac{b_r(u)}{\|u\|_n^2} \\
+& \leq 80\frac{(1 + L_pr)}{\sigma_K}\biggl(1 + A_1(\theta + \wt{\delta})\biggr) \biggl(1 + A_3\Bigl(\theta + \frac{\wt{\delta}}{r}\Bigr)\biggr) \frac{\lambda_k(G_{n,r})}{nr^{d + 2}}
 \end{align*}
@@
 
-We start by defining the density $\wt{p}_n$, which will be piecewise constant over a particular partition $\mc{Q}$ of $\Xset$; specifically if $x \in Q$ for $Q \in \mc{Q}$, then
+We start by defining the density $\wt{p}_n$, which will be piecewise constant over a particular partition $\mc{Q}$ of $\Xset$. Specifically, for each $Q$ in $\mc{Q}$ and every $x \in Q$, we set
 \begin{equation}
@@
 \end{equation}
-where $\vol(\cdot)$ denotes the full-dimensional Lebesgue measure.
+where $\vol(\cdot)$ denotes the Lebesgue measure.
 
@@
 \begin{equation*}
-a_1(d, \Xset) := \frac{1}{3\bigl(L_{\Psi} L_{\Phi}\bigr)^d}~~\textrm{and}~~A_1(d,\Xset) := \max \Bigl\{2N(\Xset),L_p L_{\Psi} L_{\Phi} \Bigr\}
+a_1 := \frac{1}{3\bigl(L_{\Psi} L_{\Phi}\bigr)^d}~~\textrm{and}~~A_1 := \max \Bigl\{2N(\Xset),L_p L_{\Psi} L_{\Phi} \Bigr\}
 \end{equation*}
@@
 
-Finally, to prove part (2) of Proposition~\ref{prop:isometry}, we require Lemma~\ref{lem:smoothening_error}, which gives an estimate on the error $\Lambda_h f - f$ in $\Leb^2(P)$ norm.
+Finally, to prove part (2) of Proposition~\ref{prop:isometry}, we require Lemma~\ref{lem:smoothening_error}, which gives an estimate on the error $\Lambda_h f - f$ in $\|\cdot\|_P^2$ norm.
 \begin{lemma}[c.f Lemma 8 of \cite{trillos2019}, Lemma 5.4 of \cite{burago2014}]
@@
 	\label{eqn:smoothening_error_norm}
-	\bigl\|\Lambda_hf\bigr\|_{\Leb^2(P)}^2 \leq \frac{2p_{\max}}{p_{\min}} \bigl\|f\bigr\|_{\Leb^2(P)}^2
+	\bigl\|\Lambda_hf\bigr\|_{P}^2 \leq \frac{2p_{\max}}{p_{\min}} \bigl\|f\bigr\|_{P}^2
 	\end{equation}
@@
 	\label{eqn:smoothening_error_energy}
-	\bigl\|\Lambda_hf - f\bigr\|_{\Leb^2(P)}^2 \leq \frac{2}{\sigma_Kp_{\min}} r E_r(f)
+	\bigl\|\Lambda_hf - f\bigr\|_{P}^2 \leq \frac{2}{\sigma_Kp_{\min}} r E_r(f)
 	\end{equation}
@@
 \begin{align*}
-\bigl\|\Lambda_rf\bigr\|_{\Leb^2(P)}^2 & \leq \frac{2}{r^d p_{\min}} \int_{\Xset} \int_{\Xset} \eta_r(x',x) \bigl[f(x')\bigr]^2 p(x') p(x) \,dx' \,dx \\ 
+\bigl\|\Lambda_rf\bigr\|_{P}^2 & \leq \frac{2}{r^d p_{\min}} \int_{\Xset} \int_{\Xset} \eta_r(x',x) \bigl[f(x')\bigr]^2 p(x') p(x) \,dx' \,dx \\ 
 & \leq \frac{2p_{\max}}{p_{\min}} \int_{\Xset} \bigl[f(x')\bigr]^2 p(x') \,dx' \\
-& = \frac{2p_{\max}}{p_{\min}} \|f\|_{\Leb^2(P)}^2.
+& = \frac{2p_{\max}}{p_{\min}} \|f\|_{P}^2.
 \end{align*}
@@
 \end{equation}
-as well as implying that the norms $\|f\|_{\Leb^2(P)}$ and $\|f\|_{\Leb^2(P_n)}$ satisfy
+as well as implying that the norms $\|f\|_{P}$ and $\|f\|_{n}$ satisfy
 \begin{equation}
 \label{eqn:calder19_2}
-\bigl(1 - A_1(\theta + \wt{\delta})\bigr) \|f\|_{\Leb^2(P)}^2 \leq \|f\|_{\Leb^2(\wt{P}_n)}^2 \leq \bigl(1 + A_1(\theta + \wt{\delta})\bigr) \|f\|_{\Leb^2(P)}^2.
+\bigl(1 - A_1(\theta + \wt{\delta})\bigr) \|f\|_{P}^2 \leq \|f\|_{\Leb^2(\wt{P}_n)}^2 \leq \bigl(1 + A_1(\theta + \wt{\delta})\bigr) \|f\|_{P}^2.
 \end{equation}
@@
 \textit{Proof of (1).}
-We begin by upper bounding $\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}$. By the Cauchy-Schwarz inequality and the bound on $\|\wt{p}_n - p\|_{\infty}$ in~\eqref{eqn:calder19_1},
+We begin by upper bounding $\bigl\|\wt{P}f\bigr\|_{n}$. By the Cauchy-Schwarz inequality and the bound on $\|\wt{p}_n - p\|_{\infty}$ in~\eqref{eqn:calder19_1},
 \begin{align*}
@@
 \label{pf:prop_isometry_1}
-\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}^2 \leq \biggl(1 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \bigl\|f\bigr\|_{\Leb^2(P)}^2.
-\end{equation}
-Now, noticing that $\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)} = \bigl\|\wt{P}^{\star}\wt{P}f\bigr\|_{\Leb^2(\wt{P}_n)}$, we can use the upper bound~\eqref{pf:prop_isometry_1} to show that
+\bigl\|\wt{P}f\bigr\|_{n}^2 \leq \biggl(1 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \bigl\|f\bigr\|_{P}^2.
+\end{equation}
+Now, noticing that $\bigl\|\wt{P}f\bigr\|_{n} = \bigl\|\wt{P}^{\star}\wt{P}f\bigr\|_{\Leb^2(\wt{P}_n)}$, we can use the upper bound~\eqref{pf:prop_isometry_1} to show that
 \begin{align}
-\Bigl|\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}^2 - \bigl\|f\bigr\|_{\Leb^2(P)}^2\Bigr| & \leq \Bigl|\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}^2 - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2\Bigr| + \Bigl|\bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2 - \bigl\|f\bigr\|_{\Leb^2(P)}^2\Bigr| \nonumber \\
-& \overset{(i)}{\leq} \Bigl|\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}^2 - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2\Bigr|  + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{\Leb^2(P)}^2 \\
-& \overset{(ii)}{\leq} \biggl(2 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \Bigl|\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)} - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}\Bigr| \cdot \bigl\|f\bigr\|_{\Leb^2(P)} + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{\Leb^2(P)}^2 \nonumber \\
-& \leq \biggl(2 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)} \cdot \bigl\|f\bigr\|_{\Leb^2(P)} + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{\Leb^2(P)}^2 \label{pf:prop_isometry_2}
+\Bigl|\bigl\|\wt{P}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{P}^2\Bigr| & \leq \Bigl|\bigl\|\wt{P}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2\Bigr| + \Bigl|\bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2 - \bigl\|f\bigr\|_{P}^2\Bigr| \nonumber \\
+& \overset{(i)}{\leq} \Bigl|\bigl\|\wt{P}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2\Bigr|  + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{P}^2 \\
+& \overset{(ii)}{\leq} \biggl(2 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \Bigl|\bigl\|\wt{P}f\bigr\|_{n} - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}\Bigr| \cdot \bigl\|f\bigr\|_{P} + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{P}^2 \nonumber \\
+& \leq \biggl(2 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)} \cdot \bigl\|f\bigr\|_{P} + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{P}^2 \label{pf:prop_isometry_2}
 \end{align}
@@
 \begin{align}
-\Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(P)}^2 - \|u\|_{\Leb^2(P_n)}^2\Bigr| & \leq \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(P)}^2 - \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2\Bigr| + \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 - \|u\|_{\Leb^2(P_n)}^2\Bigr| \nonumber \\
-& \leq A_1(\theta + \wt{\delta}) \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 + \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 - \|u\|_{\Leb^2(P_n)}^2\Bigr| \nonumber\\
-& = A_1(\theta + \wt{\delta}) \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 + \Bigl(\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} + \|u\|_{\Leb^2(P_n)}\Bigr) \cdot \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|u\|_{\Leb^2(P_n)}\Bigr| \label{pf:prop_isometry_4}
+\Bigl|\|\wt{\mc{I}}u\|_{P}^2 - \|u\|_{n}^2\Bigr| & \leq \Bigl|\|\wt{\mc{I}}u\|_{P}^2 - \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2\Bigr| + \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 - \|u\|_{n}^2\Bigr| \nonumber \\
+& \leq A_1(\theta + \wt{\delta}) \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 + \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 - \|u\|_{n}^2\Bigr| \nonumber\\
+& = A_1(\theta + \wt{\delta}) \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 + \Bigl(\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} + \|u\|_{n}\Bigr) \cdot \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|u\|_{n}\Bigr| \label{pf:prop_isometry_4}
 \end{align}
-To upper bound the second term in the above expression, we first note that~$\|u\|_{\Leb^2(P_n)} = \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}$, and thus
+To upper bound the second term in the above expression, we first note that~$\|u\|_{n} = \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}$, and thus
 \begin{align}
-\Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|u\|_{\Leb^2(P_n)} \Bigr| & = \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}\Bigr| \nonumber \\
+\Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|u\|_{n} \Bigr| & = \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}\Bigr| \nonumber \\
 & \overset{(iii)}{\leq} \|\Lambda_{\wt{r}}\wt{\mc{P}}^{\star}u - \wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)} \nonumber \\
@@
 \begin{align*}
-\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 & \leq \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \|\wt{\mc{I}}u\|_{\Leb^2(P)}^2 \\
-& \leq \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(P)}^2 \\
+\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 & \leq \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \|\wt{\mc{I}}u\|_{P}^2 \\
+& \leq \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \|\wt{\mc{P}}^{\star}u\|_{P}^2 \\
 & \leq \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)^2 \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}^2 \\
-& = \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)^2 \|u\|_{\Leb^2(P_n)}^2
+& = \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)^2 \|u\|_{n}^2
 \end{align*}
@@
 	\item 
-	\label{asmp:r_small_3} For any $x \in \Xset$, letting $\mc{B}_x:= (\Xset - x)/r~ \cap~B(0,1)$, we have that $\int_{\mc{B}_x} \psi(\|z\|) \,dz \geq a_1(\Xset,d) \cdot \int_{B(0,1)} \psi(\|z\|) \,dz$.
+	\label{asmp:r_small_3} Set $\mc{B}_x:= (\Xset - x)/r~ \cap~B(0,1)$. For any $x \in \Xset$ and any function $\psi: [0,\infty) \to [0,\infty)$,
+	\begin{equation}
+	\label{eqn:lb_expected_degree}
+	\int_{\mc{B}_x} \psi(\|z\|) \,dz \geq a_1(\Xset,d) \cdot \int_{B(0,1)} \psi(\|z\|) \,dz.
+	\end{equation}
 \end{enumerate}
@@
 \paragraph{Proof of Lemma~\ref{lem:empirical_norm_sobolev}.}
-To prove~\eqref{eqn:empirical_norm_sobolev} we will show
-\begin{equation*}
-\mathbb{E}\bigl[\norm{f}_n^4\bigr] \leq \left(1 + \frac{1}{b^2}\right)\cdot \left(\mathbb{E}\bigl[\norm{f}_n^2\bigr]\right)^2
-\end{equation*}
-whence the claim follows from the Paley-Zygmund inequality (Lemma~\ref{lem:paley_zygmund}).
+To establish~\eqref{eqn:empirical_norm_sobolev}, it is sufficient to show that
+\begin{equation*}
+\mathbb{E}\bigl[\norm{f}_n^4\bigr] \leq \left(1 + \frac{1}{b^2}\right)\cdot \left(\mathbb{E}\bigl[\norm{f}_n^2\bigr]\right)^2;
+\end{equation*}
+then \eqref{eqn:empirical_norm_sobolev} follows from the Paley-Zygmund inequality (Lemma~\ref{lem:paley_zygmund}).
 Since $p \leq p_{\max}$ is uniformly bounded, we can relate $\mathbb{E}\bigl[\norm{f}_n^4\bigr]$ to the $\Leb^4$ norm,
@@
 
-\paragraph{Proof of Lemma~\ref{lem:empirical_norm_sobolev}.}
-
-\textcolor{red}{(TODO)}
-
-\section{Proof of theorems}
+\section{Proof of Main Results}
+\label{sec:main_results}
 
@@
 \end{align*}
+\subsection{Bounds on $\Leb^2(\Xset)$ error under Lipschitz assumption.}
+Suppose $f_0$ has bounded Lipschitz constant.
+Let $V_1,\ldots,V_n$ denote the Voronoi tesselation of $\Xset$ with respect to $X_1,\ldots,X_n$.
+Extend $\wh{f}$ over $\Xset$ by taking it piecewise constant over the Voronoi cells, i.e.
+\begin{equation*}
+\wh{f}(x) := \sum_{i = 1}^{n} \cdot \wh{f}_i \1\{x \in V_i\}.
+\end{equation*}
+Note that we are abusing notation slightly by also using $\wh{f}$ to refer to this extension.
+
+In Proposition~\ref{prop:out_of_sample_error}, we establish that the out-of-sample error $\|\wh{f} - f_0\|_{\Leb^2(\Xset)}$ will not be too much larger than the in-sample error $\|\wh{f} - f_0\|_n$.
+\begin{proposition}
+	\label{prop:out_of_sample_error}
+	Suppose $f_0$ satisfies $|f_0(x') - f_0(x)| \leq M \|x' - x\|$ for all $x',x \in \mc{X}$. Then, with probability at least $1 - \delta$ it holds that
+	\begin{equation*}
+	\|\wh{f} - f_0\|_{\Leb^2(\Xset)}^2 \leq C \log(1/\delta) \biggl(\log(n)\cdot \|\wh{f} - f_0\|_n^2 + M^2\Bigl(\frac{\log n}{n}\Bigr)^{2/d}\biggr).
+	\end{equation*}
+\end{proposition}
+Note that $n^{-2/d} \ll n^{-2/(2 +d)}$.
+Therefore Proposition~\ref{prop:out_of_sample_error} together with Theorem~\ref{thm:laplacian_smoothing_estimation1} imply that with high probability, $\wh{f}$ achieves the nearly-optimal (up to a factor of $\log n$) estimation out-of-sample error---that is, $\|\wh{f} - f_0\|_{\Leb^2(\Xset)}^2 \leq C \log(n) M^{2d/(2 + d)}n^{-2/(2+d)}$---as long as $M \leq Cn^{1/d}$.
+\paragraph{Proof of Proposition~\ref{prop:out_of_sample_error}.}
+Suppose $x \in V_i$, so that we can upper bound the pointwise squared error $|\wh{f}(x) - f(x)|^2$ using the triangle inequality:
+\begin{equation*}
+\bigl(\wh{f}(x) - f_0(x)\bigr)^2 = \bigl(\wh{f}(X_i) - f_0(x)\bigr)^2 \leq 2\bigl(\wh{f}(X_i) - f_0(X_i)\bigr)^2 + 2\bigl(f_0(X_i) - f_0(x)\bigr)^2
+\end{equation*}
+Integrating both sides of the inequality, we have
+\begin{align*}
+\int_{\Xset} \bigl(\wh{f}(x) - f_0(x)\bigr)^2 \,dx & \leq 2  \sum_{i = 1}^{n} \int_{V_i} \Bigl(\wh{f}(X_i) - f_0(X_i)\Bigr)^2 \,dx + 2 \sum_{i = 1}^{n} \int_{V_i} \Bigl(f_0(X_i) - f_0(x)\Bigr)^2 \dx \\
+& = 2 \sum_{i = 1}^{n} \vol(V_i) \Bigl(\wh{f}(X_i) - f_0(X_i)\Bigr)^2 + 2 \sum_{i = 1}^{n} \int_{V_i} \Bigl(f_0(X_i) - f_0(x)\Bigr)^2 \dx
+\end{align*}
+and so by invoking the Lipschitz property of $f_0$, we obtain
+\begin{equation}
+\label{pf:out_of_sample_error_0}
+\|\wh{f} - f\|_{\Leb^2(\Xset)}^2 \leq 2 \sum_{i = 1}^{n} \vol(V_i) \Bigl(\wh{f}(X_i) - f_0(X_i)\Bigr)^2 + 2 \sum_{i = 1}^{n} \Bigl(\mathrm{diam}(V_i)\Bigr)^2
+\end{equation}
+where we write $\mathrm{diam}(V)$ for the diameter of a set $V$.
+
+Now we will use some results of \citep{chaudhuri2010} regarding uniform concentration of empirical counts. Letting
+\begin{equation*}
+\varepsilon_n := \biggl(\frac{2C_{o}\log(1/\delta)d\log n}{\nu_dp_{\min}a_0n}\biggr)^{1/d}
+\end{equation*}
+we have that for every $x \in \Xset$, $P(B(x,\varepsilon_n)) \geq 2C_{o}\log(1/\delta)d\frac{\log n}{n}$.
+Consequently, by Lemma~16 of \citep{chaudhuri2010} it holds that with probability at least $1 - \delta$,
+\begin{equation}
+\label{pf:out_of_sample_error_1}
+\textrm{for all $x \in \mc{X}$},~~ B(x,\varepsilon) \cap \{X_1,\ldots,X_n\} \neq \emptyset.
+\end{equation}
+But if~\eqref{pf:out_of_sample_error_1} is true, it must also hold that for each $i = 1,\ldots,n$ and for every $x \in V_i$, the distance $\|x - X_i\| \leq \varepsilon_n$, and so
+\begin{equation*}
+\max_{i = 1,\ldots,n} \mathrm{diam}(V_i) \leq 2\varepsilon_n.
+\end{equation*}
+Plugging back in to~\eqref{pf:out_of_sample_error_0}, and using the upper bound volume $\vol(V_i) \leq \nu_d \bigl(\mathrm{diam}(V_i)\bigr)^d$, we obtain the desired upper bound on $\|\wh{f} - f\|_{\Leb^2(\Xset)}^2$.
+
+\textcolor{red}{(TODO)}: Why do we have the lower bound on $P(B(x,\varepsilon_n))$?.
 
 \section{Concentration Inequalities}
+\label{sec:concentration}
 \begin{lemma}
@@
 
+Let $Z_1,\ldots,Z_n$ be independently distributed and bounded random variables, such that $\Ebb[Z_i] = \mu_i$.
+Let $S_n = Z_1 + \ldots + Z_n$ and $\mu = \mu_1 + \ldots + \mu_n$.
+The multiplicative form of Hoeffding's inequality gives sharp bounds when $\mu \ll 1$.
+\begin{lemma}[Hoeffding's Inequality, multiplicative form]
+	\label{lem:hoeffding}
+	Suppose $Z_i$ are independent random variables, which satisfy $Z_i \in [0,B]$ for $i = 1,\ldots,n$. For any $0 < \delta < 1$, it holds that
+	\begin{equation*}
+	\Pbb\biggl(\Bigl|S_n - \mu\Bigr| \geq \delta \mu\biggr) \leq 2\exp\biggl(-\frac{\delta^2\mu}{3B^2}\biggr)
+	\end{equation*}
+\end{lemma}
+We use Lemma~\ref{lem:hoeffding}, along with properties of the kernel $K$ and density $p$, to upper bound the maximum degree in our neighborhood graph, which we denote by $D_{\max}(G_{n,r}) := \max_{i = 1,\ldots,n} D_{ii}$.
+\begin{lemma}
+	\label{lem:max_degree}
+	Under the conditions of Lemma~\ref{lem:neighborhood_eigenvalue},
+	\begin{equation*}
+	D_{\max}(G_{n,r}) \leq 2p_{\max}nr^d,
+	\end{equation*}
+	with probability at least $1 - 2n\exp\Bigl(-nr^da_0p_{\min}/(3[K(0)]^2)\Bigr)$. 
+\end{lemma}
+\paragraph{Proof of Lemma~\ref{lem:max_degree}.}
+Fix $x \in \Xset$, and set
+\begin{equation*}
+D_{n,r}(x) :=  \sum_{i = 1}^{n} K\biggl(\frac{\|X_i - x\|}{r}\biggr);
+\end{equation*}
+note that $D_{n,r}(X_i)$ is just the degree of $X_i$ in $G_{n,r}$. By Hoeffding's inequality
+\begin{equation}
+\label{pf:max_degree_1}
+\Pbb\biggl(\Bigl|D_{n,r}(x) - \Ebb\bigl[D_{n,r}(x)\bigr]\Bigr| \geq \delta \Ebb\bigl[D_{n,r}(x)\bigr]\biggr) \leq 2\exp\biggl(-\frac{\delta^2\Ebb\bigl[D_{n,r}(x)\bigr]}{3[K(0)]^2}\biggr)
+\end{equation}
+
+Now we lower bound $\Ebb[D_{n,r}(x)]$ using the boundedness of the density $p$, and the fact that $\Xset$ has Lipschitz boundary:
+\begin{align*}
+\Ebb\bigl[D_{n,r}(x)\bigr] & = n \int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x) \,dx \\
+& \geq n p_{\min} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx \\
+& =  nr^d p_{\min}\int_{\mc{B}_x} K\bigl(\|z\|\bigr) \,dz \\
+& \geq nr^d a_0 p_{\min}
+\end{align*}
+with the final inequality following from~\eqref{eqn:lb_expected_degree}, and the normalization $\int_{\Rd} K(\|z\|) \,dz = 1$. Similar derivations yield the upper bound
+\begin{equation*}
+\Ebb\bigl[D_{n,r}(x)\bigr] \leq nr^{d} p_{\max}
+\end{equation*} 
+and plugging these bounds in to~\eqref{pf:max_degree_1}, we determine that
+\begin{equation*}
+\Pbb\biggl(D_{n,r}(x) \geq (1 + \delta) nr^d p_{\max}\biggr) \leq 2\exp\biggl(-\frac{\delta^2nr^da_0p_{\min}}{3[K(0)]^2}\biggr),
+\end{equation*}
+and applying a union bound, we get that
+\begin{equation*}
+\Pbb\biggl(\max_{i = 1,\ldots,n}D_{n,r}(X_i) \geq (1 + \delta) nr^d p_{\max}\biggr) \leq 2n\exp\biggl(-\frac{\delta^2nr^da_0p_{\min}}{3[K(0)]^2}\biggr);
+\end{equation*}
+taking $\delta = 1$ gives the claimed upper bound.
+
+\clearpage
+
+\bibliographystyle{plainnat}
+\bibliography{../../../graph_testing_bibliography}
+
 \end{document}

### 2020-10-22 notes/manuscripts/AISTATS/supplement_minimax_optimal_laplacian_smoothing.tex
Final draft of AISTATS supplement.

@@
 \documentclass[twoside]{article}
+\usepackage{aistats2021}
 \usepackage{amsmath, amsfonts, amsthm, amssymb}
@@
 \usepackage{verbatim}
-\usepackage{fullpage}
 
@@
 
+%%% New section numbering
+\renewcommand{\thesection}{A\arabic{section}}
+
 \begin{document}
 
-\title{Supplementary Materials for Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs}
-\date{}
-\maketitle
+\onecolumn	
+\aistatstitle{Supplementary Materials for ``Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs''}
 
@@
 In Section~\ref{sec:fixed_graph_error_bounds}, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to (functionals of) the graph $G$, and allow us to upper bound the error of $\wh{f}$ and $\wh{\varphi}$ conditional on the design $\mathbf{X} = x$.
-In Sections~\ref{sec:graph_sobolev_seminorm}, \ref{sec:graph_eigenvalues}, and \ref{sec:empirical_norm}, we give all our necessary probabilistic estimates on these functionals, for the particular random neighborhood graph $G = G_{n,r}$.
+In Sections~\ref{sec:graph_sobolev_seminorm}, \ref{sec:graph_eigenvalues}, \ref{sec:empirical_norm}, and \ref{sec:manifold} we give all our necessary probabilistic estimates on these functionals, for the particular random neighborhood graph $G = G_{n,r}$.
 It is in these sections where we invoke our various assumptions on the distribution $P$ and regression function $f_0$.
@@
 In Section~\ref{sec:concentration}, we state a few concentration bounds that we use repeatedly in our proofs.
+
+\vfill
+
+\clearpage
 
@@
 \begin{equation*}
-\Ebb\Bigl[f^T \Lap_{n,r} f\Bigr] \leq p_{\max}^2 \sigma_K n^2 r^{d + 2} |f|_{H^1(\Xset)}^2
+\Ebb\Bigl[f^T \Lap_{n,r} f\Bigr] \leq (1 + L_pr)^2 \sigma_K n^2 r^{d + 2} |f|_{H^1(\Xset)}^2
 \end{equation*}
@@
 \begin{equation*}
-f^T \Lap_{n,r} f = \frac{1}{2}\sum_{i, j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 \mathbf{W}_{ij},
+f^T \Lap_{n,r} f = \frac{1}{2}\sum_{i, j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 W_{ij},
 \end{equation*}
@@
 \begin{equation}
-\Ebb\biggl[\bigl(f(X') - f(X)\bigr)^2K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr] \leq p_{\max}^2 \int_{\Xset} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K_r(x',x) \,dx' \,dx.
-\label{pf:first_order_graph_sobolev_seminorm_2}
-\end{equation}
-and it remains now to bound the integral. Taking a first-order Taylor expansion of $f$, we get
+\Ebb\biggl[\bigl(f(X') - f(X)\bigr)^2K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr] = \int_{\Xset} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx, \label{pf:first_order_graph_sobolev_seminorm_2}
+\end{equation}
+so that it remains now to bound the integral. Taking a first-order Taylor expansion of $f$, we get
 \begin{align}
-\int_{\Xset} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K_r(x',x) \,dx' \,dx & = \int_{\Xset} \int_{\Xset} \biggl[\int_{0}^{1} \nabla f\bigl(x + t(x' - x)\bigr)^T (x' - x)\,dt\biggr]^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx' \,dx \nonumber \\
-& \overset{(i)}{\leq} \int_{\Xset} \int_{\Xset} \int_{0}^{1} \biggl[\nabla f\bigl(x + t(x' - x)\bigr)^T (x' - x)\biggr]^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dt\,dx' \,dx \nonumber \\
-& \overset{(ii)}{=} r^{d + 2} \int_{\Xset} \int_{\mc{Z}_r(x)} \int_{0}^{1} \biggl[\nabla f\bigl(x + trz\bigr)^T z\biggr]^2 K\bigl(\|z\|\bigr) \,dt\,dz \,dx \\
-&  \overset{(iii)}{\leq} r^{d + 2} \int_{\Xset} \int_{\Reals^d} \int_{0}^{1} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) \,dt \,dz \,d\wt{x} \nonumber \\
-& = r^{d + 2} \int_{\Xset} \int_{\Reals^d} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) \,dz \,d\wt{x} \label{pf:first_order_graph_sobolev_seminorm_3}
+\int_{\Xset} & \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \\
+& = \int_{\Xset} \int_{\Xset} \biggl[\int_{0}^{1} \nabla f\bigl(x + t(x' - x)\bigr)^T (x' - x)\,dt\biggr]^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \nonumber \\
+& \overset{(i)}{\leq} \int_{\Xset} \int_{\Xset} \int_{0}^{1} \biggl[\nabla f\bigl(x + t(x' - x)\bigr)^T (x' - x)\biggr]^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dt\,dx' \,dx \nonumber \\
+& \overset{(ii)}{=} r^{d + 2} \int_{\Xset} \int_{\mc{Z}_r(x)} \int_{0}^{1} \biggl[\nabla f\bigl(x + trz\bigr)^T z\biggr]^2 K\bigl(\|z\|\bigr) p(zr + x) p(x) \,dt\,dz \,dx \\
+&  \overset{(iii)}{\leq} r^{d + 2} \int_{\Xset} \int_{\Reals^d} \int_{0}^{1} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) p\bigl(\wt{x} + [1-t]rz\bigr) p\bigl(\wt{x} + trz\bigr) \,dt \,dz \,d\wt{x} \nonumber \\
+& \overset{(iv)}\leq \biggl(1 + \frac{L_pr}{p_{\min}}\biggr)^2 r^{d + 2} \int_{\Xset} \int_{\Reals^d} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) p(\wt{x})^2 \,dz \,d\wt{x} \label{pf:first_order_graph_sobolev_seminorm_3}
 \end{align}
-where $(i)$ follows by Jensen's inequality, $(ii)$ follows by substituting $z = (x' - x)/r$ with $\mc{Z}_r(x) := \{z: zr + x \in \Xset\}$ and $(iii)$ by substituting $\wt{x} = x + trz$ and noting that
-\begin{equation*}
-x \in \Xset ~~\textrm{and}~~ z \in \mc{Z}_r(x) \Longrightarrow x + trz \in \Xset ~~\textrm{and}~~ z \in \Reals^d.
-\end{equation*}
+where $(i)$ follows by Jensen's inequality, $(ii)$ follows by substituting $z = (x' - x)/r$ with $\mc{Z}_r(x) := \{z: zr + x \in \Xset\}$ $(iii)$ by substituting $\wt{x} = x + trz$ and noting that
+\begin{equation*}
+x \in \Xset ~~\textrm{and}~~ z \in \mc{Z}_r(x) \Longrightarrow x + trz \in \Xset ~~\textrm{and}~~ z \in \Reals^d,
+\end{equation*}
+and $(iv)$ by the Lipschitz property of $p$.
+
 Now, writing $\bigl(\nabla f(\wt{x}) ^T z\bigr)^2 = \bigl(\sum_{i = 1}^{d} z_{i} f^{(e_i)}(x) \bigr)^2$, expanding the square and integrating, we have that for any $\wt{x} \in \Xset$,
@@
 \end{equation*}
-To avoid confusion, in this section we write $\lambda_k(G_{n,r})$ for the $k$th smallest eigenvalue of the graph Laplacian matrix $\Lap_{n,r}$ and $\lambda_k(\Delta_P)$ for the $k$th smallest eigenvalue of $\Delta_P$ \footnote{Under the assumptions}.
-Some other notation: throughout this section, we will write $A, A_1,\ldots$ and $a,a_1,\ldots$ for constants which do not depend on $n$, and we let $\theta$ and $\wt{\delta}$ denote (small) positive numbers, which always satisfy the following restrictions.
+To avoid confusion, in this section we write $\lambda_k(G_{n,r})$ for the $k$th smallest eigenvalue of the graph Laplacian matrix $\Lap_{n,r}$ and $\lambda_k(\Delta_P)$ for the $k$th smallest eigenvalue of $\Delta_P$ \footnote{Under the assumptions~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density}, the operator $\Delta_p$ has a discrete spectrum; see \cite{garciatrillos18} for more details.}.
+Some other notation: throughout this section, we will write $A, A_0, A_1,\ldots$ and $a,a_0,a_1,\ldots$ for constants which may depend on $\Xset$, $d$, $K$, and $p$ but do not depend on $n$.
+We let $L_K$ denote the Lipschitz constant of the kernel $K$, and we let $\theta$ and $\wt{\delta}$ denote (small) positive numbers, which always satisfy the following restrictions.
 \begin{enumerate}[label=(S\arabic*)]
 	\item 
-	\begin{equation*}
-	\theta + \wt{\delta} \leq \frac{1}{4A_1}~~\textrm{and}~~C_0\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq \wt{\delta} \leq r
+	\label{asmp:smallness}
+	\begin{equation*}
+	\theta + \wt{\delta} \leq \frac{1}{4\max\{A_1,A\}}~~\textrm{and}~~n^{-1/d} \leq \wt{\delta} \leq \min\biggl\{\frac{1}{2^{d + 3}A_0}, \frac{1}{A_3}, \frac{K(1)}{8L_KA_0}\biggr\}r
 	\end{equation*}
@@
 	\end{equation}
-	with probability at least $1 - A_1n\exp(-a_1n\theta^2\wt{\delta}^{m})$, it holds that
+	with probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{m})$, it holds that
 	\begin{equation}
@@
 \begin{equation*}
-\ell_{\star} = \floor{\biggl(\frac{\bigl(1/(2A_1) - (\theta + \wt{\delta})\bigr)}{rA_2^{1/2}}\biggr)^d},
+\ell_{\star} = \floor{\biggl(\frac{\bigl(1/(2A) - (\theta + \wt{\delta})\bigr)}{rA_2^{1/2}}\biggr)^d},
 \end{equation*}
@@
 \begin{align*}
-\sqrt{\lambda_{\ell_{\star}}(\Delta)} \leq \frac{1}{r}\biggl(\frac{1}{2A_1} - (\theta + \wt{\delta})\biggr) 
+\sqrt{\lambda_{\ell_{\star}}(\Delta_P)} \leq \frac{1}{r}\biggl(\frac{1}{2A} - (\theta + \wt{\delta})\biggr) 
 \end{align*}
 so that condition~\eqref{eqn:neighborhood_eigenvalue_1} is satisfied.
-We will therefore assume the conclusions of Theorem~\ref{thm:neighborhood_eigenvalue}.
+We may therefore assume the conclusions of Theorem~\ref{thm:neighborhood_eigenvalue}.
 Along with~\eqref{eqn:weyls_law}, this implies the following bounds on the graph Laplacian eigenvalues:
 \begin{equation*}
-\frac{a_2}{A_1} nr^{d + 2} k^{2/d} \leq \lambda_k(G_{n,r}) \leq \frac{A_2}{a_1} nr^{d + 2} k^{2/d}~~\textrm{for all $2 \leq k \leq \ell_{\star}$}.
+\frac{a}{A_2} nr^{d + 2} k^{2/d} \leq \lambda_k(G_{n,r}) \leq \frac{A}{a_2} nr^{d + 2} k^{2/d}~~\textrm{for all $2 \leq k \leq \ell_{\star}$}.
 \end{equation*}
@@
 \lambda_k(G_{n,r}) & \geq \lambda_{\ell_{\star}}(G_{n,r}) \\
-& \geq \frac{a_2}{A_1}nr^{d + 2}\ell_{\star}^{2/d} \\
-& = \frac{a_2}{A_1}nr^{d + 2} \floor{\biggl(\frac{\bigl(1/(2A_1) - (\theta + \wt{\delta})\bigr)}{rA_2^{1/2}}\biggr)^d}^2 \\
-& \geq \frac{a_2}{32A_1^3 A_2} nr^{d}
-\end{align*}
-with the last line following since $\theta + \wt{\delta} \leq \frac{1}{4A_1}$.
+& \geq \frac{a_2}{A}nr^{d + 2}\ell_{\star}^{2/d} \\
+& = \frac{a_2}{A}nr^{d + 2} \floor{\biggl(\frac{\bigl(1/(2A) - (\theta + \wt{\delta})\bigr)}{rA_2^{1/2}}\biggr)^d}^{2/d} \\
+& \geq \frac{a_2}{32A^3 A_2} nr^{d}
+\end{align*}
+with the last line following since $\theta + \wt{\delta} \leq \frac{1}{4A}$.
 On the other hand, for any graph $G$ the maximum eigenvalue of the Laplacian is upper bounded by twice the maximum degree \citep{chung97}.
@@
 \end{equation*}
-with probability at least $1 - 2n\exp\Bigl(-nr^dp_{\min}/(3K(0)^2)\Bigr)$. Thus we have established each of the inequalities in Lemma~\ref{lem:neighborhood_eigenvalue}.
+with probability at least $1 - 2n\exp\Bigl(-nr^dp_{\min}/(3K(0)^2)\Bigr)$.
+Setting $\theta$ and $\wt{\delta}$ as large as possible while ensuring that~\ref{asmp:smallness} and~\eqref{eqn:neighborhood_eigenvalue_1} are satisfied---so that $\theta \asymp 1$ and $\wt{\delta} \asymp r$---we have established that each of the inequalities in Lemma~\ref{lem:neighborhood_eigenvalue} hold with probability at least $1 - Cn\exp(-cnr^d)$.
 
@@
 We recall that for a function $u: \{X_1,\ldots,X_n\} \to \Reals$, the empirical norm is defined as $\|u\|_n^2 := \frac{1}{n} \sum_{i = 1}^{n} (u(X_i))^2$, and we let the class $\Leb^2(P_n)$ consist of those $u \in \Reals^n$ for which $\|u\|_{n} < \infty$.
-For a function $f: \Xset \to \Reals$, the continuum limit of $\|f\|_n^2$ is simply
-\begin{equation*}
-\|f\|_{P}^2 := \int_{\Xset} \bigl|f(x)\bigr|^2 p(x) \,dx,
-\end{equation*} 
-and we let $\Leb^2(P)$ be the set of functions for which $\|f\|_P < \infty$.
+For a function $f: \Xset \to \Reals$, the continuum limit of $\|f\|_n^2$ is
+\begin{equation*}
+\|f\|_{P}^2 := \int_{\Xset} \bigl|f(x)\bigr|^2 p(x) \,dx.
+\end{equation*}
 
@@
 \begin{enumerate}[(1)]
-	\item an upper bound of $b_r(u)$ by $D_2\bigl(\mc{I}(u)\bigr)$ for an appropriate choice of interpolating map $\mc{I}: \Leb^2(P_n) \to \Leb^2(\mc{X})$, and vice versa upper bound $D_2(f)$ by $b_r(\mc{P}(f))$ for an appropriate choice of discretization map $\mc{P}: \Leb^2(\mc{X}) \to \Leb^2(P_n)$,
+	\item an upper bound of $b_r(u)$ by $D_2\bigl(\mc{I}(u)\bigr)$ for an appropriate choice of interpolating map $\mc{I}: \Leb^2(P_n) \to \Leb^2(\mc{X})$, and vice versa an upper bound of $D_2(f)$ by $b_r(\mc{P}(f))$ for an appropriate choice of discretization map $\mc{P}: \Leb^2(\mc{X}) \to \Leb^2(P_n)$,
 	\item that $\mc{I}$ and $\mc{P}$ were near-isometries, meaning $\|\mc{I}(u)\|_{P} \approx \|u\|_{n}$ and $\|\mc{P}(f)\|_{P} \approx \|f\|_{n}$.
@@
 
-We will momentarily define maps $\wt{\mc{I}}$ and $\wt{\mc{P}}$ that satisfy both (1) and (2).
-These maps will be defined with respect to a particular probability measure $\wt{P}_n$, that with high probability is close in transportation distance to both $P_n$ and $P$.
-This estimate on the transportation distance---which we now give--- will be the workhorse that allows us to relate the random energy $b_r$ to the non-random $D_2$.
+We will momentarily define particular maps $\wt{\mc{I}}$ and $\wt{\mc{P}}$, and establish that they satisfy both (1) and (2).
+In order to define these maps, we must first introduce a particular probability measure $\wt{P}_n$ that with high probability is close in transportation distance to both $P_n$ and $P$.
+This estimate on the transportation distance---which we now give--- will be the workhorse that allows us to relate $b_r$ to $D_2$, and $\|\cdot\|_n$ to $\|\cdot\|_P$.
 
 \paragraph{Transportation distance between $P_n$ and $P$.}
-For a map $T$ defined on $\Xset$ and a measure $\mu$, let $T_{\sharp}\mu$ denote the push-forward of $\mu$ by $T$, i.e the measure for which
+For a measure $\mu$ defined on $\Xset$ and map $T: \Xset \to \Xset$, let $T_{\sharp}\mu$ denote the \emph{push-forward} of $\mu$ by $T$, i.e the measure for which
 \begin{equation*}
@@
 The  $\infty$-transportation distance between $\mu$ and $P_n$ is then
-\begin{equation*}
-d_{\infty}(\mu,P_n) := \inf_{T: T_{\sharp} \mu = P_n} \biggl\{\sup_{x \in \Xset}~\bigl|T(x) - x\bigr|\biggr\}
-\end{equation*}
+\begin{equation}
+\label{eqn:optimal_transport}
+d_{\infty}(\mu,P_n) := \inf_{T: T_{\sharp} \mu = P_n} \|T - \mathrm{Id}\|_{L^{\infty}(\mu)}
+\end{equation}
+where $\mathrm{Id}(x) = x$ is the identity mapping.
+
 \cite{calder2019} take $\Xset$ to be a smooth submanifold of $\Rd$ without boundary, i.e.
@@
 	\label{prop:optimal_transport}
-	Suppose $p$ satisfies~\ref{asmp:bounded_lipschitz_density} and $\Xset$ satisfies~\ref{asmp:domain}.
-Then with probability at least $1 - A_1 n \exp\bigl\{-a_1 n\theta^2\wt{\delta}^d\bigr\}$, the following statement holds: there exists a probability measure $\wt{P}_n$ with density $\wt{p}_n$ such that:
-	\begin{equation*}
-	d_{\infty}(\wt{P}_n, P_n) \leq A_1 \wt{\delta}
-	\end{equation*}
+	Suppose $\Xset$ satisfies~\ref{asmp:domain} and $p$ satisfies~\ref{asmp:bounded_lipschitz_density}.
+Then with probability at least $1 - A_0 n \exp\bigl\{-a_0 n\theta^2\wt{\delta}^d\bigr\}$, the following statement holds: there exists a probability measure $\wt{P}_n$ with density $\wt{p}_n$ such that:
+	\begin{equation}
+	\label{eqn:optimal_transport_1}
+	d_{\infty}(\wt{P}_n, P_n) \leq A_0 \wt{\delta}
+	\end{equation}
 	and
-	\begin{equation*}
-	\|\wt{p}_n - p\|_{\infty} \leq A_1\bigl(\wt{\delta} + \theta\bigr)
-	\end{equation*}
+	\begin{equation}
+	\label{eqn:optimal_transport_2}
+	\|\wt{p}_n - p\|_{\infty} \leq A_0\bigl(\wt{\delta} + \theta\bigr)
+	\end{equation}
 \end{proposition}
-For the rest of this section, we let $\wt{P}_n$ be a probability measure with density $\wt{p}_n$, that satisfies the conclusions of Proposition~\ref{prop:isometry}.
-Additionally we denote by $\wt{T}_n$ an \emph{optimal transport map} between $\wt{P}_n$ and $P_n$.
+For the rest of this section, we let $\wt{P}_n$ be a probability measure with density $\wt{p}_n$, that satisfies the conclusions of Proposition~\ref{prop:optimal_transport}.
+Additionally we denote by $\wt{T}_n$ an \emph{optimal transport map} between $\wt{P}_n$ and $P_n$, meaning a transportation map which achieves the infimum in~\eqref{eqn:optimal_transport}.
 Finally, we write $U_1,\ldots,U_n$ for the preimages of $X_1,\ldots,X_n$ under $\wt{T}_n$.
@@
 \paragraph{Interpolation and discretization maps.}
-
 The discretization map  $\wt{\mathcal{P}}: \Leb^2(\Xset) \to \Leb^2(P_n)$ is given by
@@
 \end{equation*}
-On the other hand, the interpolation map $\wt{\mc{I}}: \Leb^2(P_n) \to \Leb^2(\Xset)$ is defined as $\wt{\mc{I}}u := \Lambda_{r - 2\wt{\delta}}(\wt{\mc{P}}^{\star}u)$.
+On the other hand, the interpolation map $\wt{\mc{I}}: \Leb^2(P_n) \to \Leb^2(\Xset)$ is defined as $\wt{\mc{I}}u := \Lambda_{r - 2A_0\wt{\delta}}(\wt{\mc{P}}^{\star}u)$.
 Here, $\wt{\mc{P}}^{\star} = u \circ \wt{T}$ is the adjoint of $\wt{\mc{P}}$, i.e.
@@
 \end{equation*} 
-and $\Lambda_r$ is a smoothening of this extension, formally
-\begin{equation*}
-\Lambda_r(f) := \frac{1}{r^d\tau(x)}\int_{\Xset} \eta_r(x',x) f(x') \,dx',~~ \eta_r(x',x) := \psi\biggl(\frac{\|x' - x\|}{r}\biggr)
-\end{equation*}
-where $\psi(t) := (1/\sigma_K)\int_{t}^{\infty} s K(s) \,ds$ and $\tau(x) := (1/r^d)\int_{\Xset} \eta_r(x',x) \,dx'$ is a normalizing constant.
+and $\Lambda_{r - 2A_0\wt{\delta}}$ is a smoothening operator. To be precise, for any $h > 0$,
+\begin{equation*}
+\Lambda_h(f) := \frac{1}{h^d\tau_h(x)}\int_{\Xset} \eta_h(x',x) f(x') \,dx',~~ \eta_h(x',x) := \psi\biggl(\frac{\|x' - x\|}{r}\biggr)
+\end{equation*}
+where $\psi(t) := (1/\sigma_K)\int_{t}^{\infty} s K(s) \,ds$ and $\tau_h(x) := (1/h^d)\int_{\Xset} \eta_h(x',x) \,dx'$ is a normalizing constant.
 
@@
 	\label{prop:dirichlet_energies}
-	With probability at least $1 - A_1n\exp(-a_1n\theta^2\wt{\delta}^{m})$, we have the following.
+	With probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{m})$, we have the following.
 	\begin{enumerate}[(1)]
@@
 		\begin{equation*}
-		\sigma_{K} D_2(\wt{\mc{I}}u) \leq \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) b_r(u)
+		\sigma_{K} D_2(\wt{\mc{I}}u) \leq A_8 \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) b_r(u)
 		\end{equation*}
@@
 		\begin{equation*}
-		b_r(\wt{\mc{P}}f) \leq \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + L_p(2r + 4\wt{\delta})\Bigr) \cdot \Bigl(1 + A_4\frac{\wt{\delta}}{r}\Bigr) \sigma_{K} D_2(f).
+		b_r(\wt{\mc{P}}f) \leq \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + \frac{L_p(r + 2A_0\wt{\delta})}{p_{\min}}\Bigr)^2 \cdot \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr) \cdot \sigma_{K} D_2(f)
 		\end{equation*}
@@
 	\label{prop:isometry}
-	With probability at least $1 - A_1n\exp(-a_1n\theta^2\wt{\delta}^{m})$, we have the following.
+	With probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{m})$, we have the following.
 	\begin{enumerate}[(1)]
@@
 		\begin{equation*}
-		\Bigl|\|f\|_{P}^2 - \|\wt{P}f\|_{n}^2\Bigr| \leq A_5 \wt{\delta} \|f\|_{P} \sqrt{D_2(f)} + \frac{A_1\bigl(\theta + \wt{\delta}\bigr)}{p_{\min}} \|f\|_{P}^2
+		\Bigl|\|f\|_{P}^2 - \|\wt{P}f\|_{n}^2\Bigr| \leq A_5 r \|f\|_{P} \sqrt{D_2(f)} + A_1\bigl(\theta + \wt{\delta}\bigr)\|f\|_{P}^2
 		\end{equation*}
@@
 
-Given these propositions, the proof of Theorem~\ref{thm:neighborhood_eigenvalue} follows by the Courant-Fischer Theorem, as we now show.
+We will devote most of the rest of this section to the proofs of Propositions~\ref{prop:optimal_transport},~\ref{prop:dirichlet_energies}, and~\ref{prop:isometry}.
+First, however, we use these propositions to prove Theorem~\ref{thm:neighborhood_eigenvalue}.
 
 \paragraph{Proof of Theorem~\ref{thm:neighborhood_eigenvalue}.}
+Let $k$ be any number in $1,\ldots,\ell$.
 We start with the upper bound, proceeding exactly as in Proposition 4.4 of \citep{burago2014}.
 Let $f_1,\ldots,f_{k}$ denote the first $k$ eigenfunctions of $\Delta_P$ and set $W := \mathrm{span}(f_1,\ldots,f_k)$, so that by the Courant-Fischer principle $D_2(f) \leq \lambda_k(\Delta_P) \|f\|_{P}^2$ for every $f \in W$.
-As a result, by Part (1) of Proposition~\ref{prop:isometry} we have that for any $f \in W$
-\begin{equation*}
-\bigl\|\wt{\mc{P}}f\bigr\|_{n}^2 \geq \biggl(1 - A_5\wt{\delta} \sqrt{\lambda_{k}(\Delta_P)} - \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr)\|f\|_{P}^2  \geq \frac{1}{2} \|f\|_{P}^2,
-\end{equation*}
-where the second inequality follows by assumption~\eqref{eqn:neighborhood_eigenvalue_1}.
+As a result, by Part (1) of Proposition~\ref{prop:isometry} we have that for any $f \in W$,
+\begin{equation*}
+\bigl\|\wt{\mc{P}}f\bigr\|_{n}^2 \geq \biggl(1 - A_5r \sqrt{\lambda_{k}(\Delta_P)} - A_1(\theta + \wt{\delta})\biggr)\|f\|_{P}^2  \geq \frac{1}{2} \|f\|_{P}^2,
+\end{equation*}
+where the second inequality follows by assumption~\eqref{eqn:neighborhood_eigenvalue_1}, upon appropriate choice of $A$.
 
@@
 & = \max_{\substack{f \in W \\ f \neq 0} } \frac{b_r(\wt{\mc{P}}f)}{\bigl\|\wt{\mc{P}}f\bigr\|_{n}^2} \\
-& \leq 2\Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + L_p(2r + 4\wt{\delta})\Bigr) \cdot \Bigl(1 + A_4\frac{\wt{\delta}}{r}\Bigr)\sigma_K \lambda_{k}(\Delta_P),
+& \leq 2\Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + \frac{L_p(r + 2A_0\wt{\delta})}{p_{\min}}\Bigr)^2 \cdot \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr)\sigma_K \lambda_{k}(\Delta_P),
 \end{align*}
@@
 
-The upper bounds follows from essentially parallel reasoning.
-Recalling that $v_1,\ldots,v_k$ denote the first $k$ eigenvecotrs of $\Lap_{n,r}$, let $U := \mathrm{span}(v_1,\ldots,v_k)$, so that $nr^{d + 2} b_r(u) \leq \lambda_k(G_{n,r}) \|u\|_n^2$.
+The upper bound follows from essentially parallel reasoning.
+Recalling that $v_1,\ldots,v_k$ denote the first $k$ eigenvectors of $\Lap_{n,r}$, set $U := \mathrm{span}(v_1,\ldots,v_k)$, so that $nr^{d + 2} b_r(u) \leq \lambda_k(G_{n,r}) \|u\|_n^2$.
 By Proposition~\ref{prop:isometry}, Part (2), for every $u \in U$ it holds that
@@
 \end{align*}
-where the second to last inequality follows from the lower bound $a \lambda_k(G_{n,r}) \leq nr^{d + 2}\lambda_k(\Delta_P)$ that we just derived, and the last inequality from assumption~\eqref{eqn:neighborhood_eigenvalue_1}.
+where the second to last inequality follows from the lower bound $a \lambda_k(G_{n,r}) \leq nr^{d + 2}\lambda_k(\Delta_P)$ that we just derived, and the last inequality from assumption~\eqref{eqn:neighborhood_eigenvalue_1}, upon appropriate choice of $A$.
 
@@
 \lambda_k(\Delta_P) & \leq \max_{u \in U} \frac{D_2(\wt{\mc{I}}u)}{\|u\|_P^2} \\
-& \leq 80\frac{(1 + L_pr)}{\sigma_K}\biggl(1 + A_1(\theta + \wt{\delta})\biggr) \biggl(1 + A_3\Bigl(\theta + \frac{\wt{\delta}}{r}\Bigr)\biggr)\max_{u \in U} \frac{b_r(u)}{\|u\|_n^2} \\
-& \leq 80\frac{(1 + L_pr)}{\sigma_K}\biggl(1 + A_1(\theta + \wt{\delta})\biggr) \biggl(1 + A_3\Bigl(\theta + \frac{\wt{\delta}}{r}\Bigr)\biggr) \frac{\lambda_k(G_{n,r})}{nr^{d + 2}}
-\end{align*}
-establishing the upper bound in~\eqref{eqn:eigenvalue_bound} upon proper choice of $A$.
+& \leq 2A_8\biggl(1 + A_1(\theta + \wt{\delta})\biggr) \biggl(1 + A_3 \frac{\wt{\delta}}{r}\biggr)\max_{u \in U} \frac{b_r(u)}{\|u\|_n^2} \\
+& \leq 2A_8\biggl(1 + A_1(\theta + \wt{\delta})\biggr) \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) \frac{\lambda_k(G_{n,r})}{nr^{d + 2}}
+\end{align*}
+establishing the upper bound in~\eqref{eqn:eigenvalue_bound} when $A$ is properly chosen.
 
@@
 The rest of this section will be devoted to proving Propositions~\ref{prop:optimal_transport},~\ref{prop:dirichlet_energies} and~\ref{prop:isometry}.
-The proof of Proposition~\ref{prop:optimal_transport} is given in Subsection~\ref{subsec:proof_proposition_optimal_transport}.
 To prove the latter two propositions, it will help to introduce the intermediate energies
@@
 
+The proof of Proposition~\ref{prop:optimal_transport} is given in Subsection~\ref{subsec:proof_proposition_optimal_transport}.
 In Subsection~\ref{subsec:integrals}, we establish relationships between the (non-random) functionals $E_r(f)$ and $D_2(f)$, as well as providing estimates on some assorted integrals.
@@
 \end{equation}
-where $\vol(\cdot)$ denotes the Lebesgue measure.
+where $\vol(\cdot)$ denotes the Lebesgue measure. Then $\wt{P}_n(U) = \int_{U} \wt{p}_n(x) \dx$.
 
@@
 	where $Q_k = \wt{\delta}\Bigl([k_1 - 1,k_1] \otimes \cdots \otimes [k_d - 1,k_d]\Bigr)$ and we assume without loss of generality that $\wt{\delta}^{-1} \in \mathbb{N}$.
-	\item If $\mc{\Xset}$ is an open, connected set with smooth boundary, then by Proposition 3.2 of \textcolor{red}{GarciaTrillos14}, there exist a finite number $N(\Xset) \in \mathbb{N}$ of disjoint polytopes which cover $\Xset$.
+	\item If $\mc{\Xset}$ is an open, connected set with smooth boundary, then by Proposition 3.2 of \cite{trillos2015}, there exist a finite number $N(\Xset) \in \mathbb{N}$ of disjoint polytopes which cover $\Xset$.
 Moreover, letting $U_j$ denote the intersection of the $j$th of these polytopes with $\wb{\Xset}$, this proposition establishes that for each $j$ there exists a bi-Lipschitz homeomorphism $\Phi_j: U_j \to [0,1]^d$.
@@
 \end{itemize}
-Let us record a few facts which hold for all $\wt{Q}_{j,k} \in \mc{Q}$, which follow from the bi-Lipschitz properties of $\Phi_j$ and $\Psi$: first that
+Let us record a few facts which hold for all $\wt{Q}_{j,k} \in \mc{Q}$, and which follow from the bi-Lipschitz properties of $\Phi_j$ and $\Psi$: first that
 \begin{equation}
@@
 We now use these facts to show that $\wt{P}_n$ satisfies the claims of Proposition~\ref{prop:optimal_transport}.
-One on the one hand, by the construction~\eqref{pf:prop_optimal_transport_0} clearly there exists a transport map $\wt{T}: \Xset \to \mathbf{X}$ for which every $x \in \wt{Q}_{j,k}$ is moved to a design point $X \in {\bf X} \cap \wt{Q}_{j,k}$, and so by~\eqref{pf:prop_optimal_transport_1}
-\begin{equation*}
-\sup_{x \in \Xset} \abs{\wt{T}(x) - x} \leq  L_{\Psi} \L_{\Phi} \wt{\delta}.
+On the one hand for every $Q \in \mc{Q}$, letting $N(Q)$ denote the number of design points $\{X_1,\ldots,X_k\}$ which fall in $Q$, we have
+\begin{equation*}
+\wt{P}_n(Q) = \int_{Q} \wt{p}_n(x) \,dx = P_n(Q) = \frac{N(Q)}{n}.
+\end{equation*}
+Moreover, ignoring those cells for which $N(Q) = 0$ (since $\wt{P}_n(Q) = 0$ for such $Q$, and so they do not contribute to the essential supremum in~\eqref{eqn:optimal_transport}), appropriately dividing each remaining cell $Q \in \mc{Q}$ into $N(Q)$ subsets $S_1,\ldots,S_{N(Q)}$ of equal volume, and mapping each $S_{\ell}$ to a different design point $X_i \in Q$, we can exhibit a transport map $T$ from $\wt{P}_n$ to $P_n$ for which
+\begin{equation*}
+\|T - \mathrm{Id}\|_{L^{\infty}(\wt{P}_n)} \leq \max_{Q \in \mc{Q}} \diam(Q) \leq   L_{\Psi} \L_{\Phi} \wt{\delta}.
 \end{equation*}
@@
 \begin{equation*}
-a_1 := \frac{1}{3\bigl(L_{\Psi} L_{\Phi}\bigr)^d}~~\textrm{and}~~A_1 := \max \Bigl\{2N(\Xset),L_p L_{\Psi} L_{\Phi} \Bigr\}
+a_0 := \frac{1}{3\bigl(L_{\Psi} L_{\Phi}\bigr)^d}~~\textrm{and}~~A_0 := \max \Bigl\{2N(\Xset),L_p L_{\Psi} L_{\Phi}, L_{\Psi} L_{\Phi}\Bigr\}
 \end{equation*}
@@
 \label{subsec:integrals}
-
-Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected} relates the non-local energy $E_r(f)$ to $D_2(f)$.
-It follows in large part from the proofs of Lemma 13 of \cite{trillos2019} and Lemma 6 of \cite{burago2014}, with minor modifications to account for the boundary of $\Xset$.
-\begin{lemma}[c.f. Lemma 13 of \cite{trillos2019}, Lemma 6 of \cite{burago2014}]
-	\label{lem:first_order_graph_sobolev_seminorm_expected}
-	For any $f \in \Leb^2(\Xset)$,
-	\begin{equation*}
-	E_r(f) \leq (1 + L_pr)^2 \cdot \sigma_K D_2(f)
-	\end{equation*}
-\end{lemma}
-
-In Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, we establish the reverse of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected}. 
+Let us start by making the following observation, which we make use of repeatedly in this section.
+Let $\eta: [0,\infty) \to [0,\infty)$ be an otherwise arbitrary function.
+As a consequence of ~\ref{asmp:domain}, there exist constants $c_0$ and $a_3$ which depend on $\Xset$, such that for any $0 < \varepsilon \leq c_0$ it holds that
+\begin{equation}
+\label{eqn:integral_boundary}
+\int_{B(x,\varepsilon) \cap \Xset} \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \geq a_3 \cdot \int_{B(x,\varepsilon)} \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx'
+\end{equation}
+As a special case: when $\eta(x) = 1$, this implies $\vol\bigl(B(x,\varepsilon) \cap \Xset\bigr) \geq a_3 \nu_d \varepsilon^d$ for any $0 \varepsilon \leq c_0$.
+
+We have already upper bounded $E_r(f)$ by (a constant times) $D_2(f)$ in the proof of Lemma~\ref{lem:graph_sobolev_seminorm}.
+In Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, we establish the reverse inequality.
 \begin{lemma}[c.f. Lemma 9 of \cite{trillos2019}, Lemma 5.5 of \cite{burago2014}]
 	\label{lem:first_order_graph_sobolev_seminorm_expected_lb}
-	For any $f \in \Leb^2(\Xset)$,
-	\begin{equation*}
-	\sigma_KD_2(\Lambda_rf) \leq \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K}E_r(f)
+	For any $f \in \Leb^2(\Xset)$, and any $0 < h \leq c_0$, it holds that
+	\begin{equation*}
+	\sigma_KD_2(\Lambda_hf) \leq A_8 E_h(f).
 	\end{equation*}
@@
 
-To prove Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, we require upper and lower bounds on $\tau(x)$, as well as an upper bound on the gradient of $\tau$.
-The lower bound here---$\tau(x) \geq 1/2$---is quite a bit a looser than in the case where $\Xset$ has no boundary---$\tau(x) \geq (1 + Cr^2)^{-1}$.
-The same is the case regarding the upper bound of the size of the gradient $\|\nabla \tau(x)\|$.
+To prove Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, we require upper and lower bounds on $\tau_h(x)$, as well as an upper bound on the gradient of $\tau_h$.
+The lower bound here---$\tau_h(x) \geq 1/2$---is quite a bit a looser than what can be shown when $\Xset$ has no boundary.
+The same is the case regarding the upper bound of the size of the gradient $\|\nabla \tau_h(x)\|$.
 However, the bounds as stated here will be sufficient for our purposes.
@@
 	\label{lem:tau_bound}
-	Suppose $r$ satisfies~\ref{asmp:r_small_3}. For all $x \in \Xset$,
-	\begin{equation*}
-	\frac{1}{2} \leq \tau(x) \leq 1,
-	\end{equation*}
-	and additionally $\|\nabla \tau(x)\| \leq \frac{1}{\sigma_K r}$.
+	For any $0 < h \leq c_0$, for all $x \in \Xset$ it holds that
+	\begin{equation*}
+	a_3 \leq \tau_h(x) \leq 1.
+	\end{equation*}
+	and 
+	\begin{equation*}
+	\|\nabla \tau_h(x)\| \leq \frac{1}{\sqrt{d\sigma_K} h}.
+	\end{equation*}
 \end{lemma}
@@
 	\label{lem:smoothening_error}
-	For any $h > 0$, 
+	For any $0 < h \leq c_0$, 
 	\begin{equation}
 	\label{eqn:smoothening_error_norm}
-	\bigl\|\Lambda_hf\bigr\|_{P}^2 \leq \frac{2p_{\max}}{p_{\min}} \bigl\|f\bigr\|_{P}^2
+	\bigl\|\Lambda_hf\bigr\|_{P}^2 \leq \frac{p_{\max}}{a_3p_{\min}} \bigl\|f\bigr\|_{P}^2
 	\end{equation}
@@
 	\label{eqn:smoothening_error_energy}
-	\bigl\|\Lambda_hf - f\bigr\|_{P}^2 \leq \frac{2}{\sigma_Kp_{\min}} r E_r(f)
+	\bigl\|\Lambda_hf - f\bigr\|_{P}^2 \leq \frac{1}{a_3\sigma_Kp_{\min}} h^2 E_h(f)
 	\end{equation}
@@
 
-\paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected}.}
-Suppose $f \in C^{\infty}(\wb{\Xset})$, which we may do without loss of generality due to the density of $C^{\infty}(\wb{\Xset})$ in  $\Leb^2(\Xset)$.
-
-Taylor expanding $f(x')$ about $x' = x$, applying first Cauchy-Schwarz' and then Jensen's inequality, and then using the Lipschitz property of $p$, we have
-\begin{align*}
-E_r(f) & = \frac{1}{r^{d + 2}}\int_{\Xset} \int_{\Xset} (f(x') - f(x))^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \\
-& = \frac{1}{r^{d + 2}}\int_{\Xset} \int_{\Xset} \biggl(\sum_{i = 1}^{d} (x' - x)^{e_i} \int_0^1 f^{(e_i)}\bigl(x + t(x' - x)\bigr) \,dt\biggr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \\
-& \leq \frac{1}{r^{d + 2}}\int_{\Xset} \int_{\Xset} \|x' - x\|^2 \sum_{i = 1}^{d}\biggl(\int_0^1 f^{(e_i)}\bigl(x + t(x' - x)\bigr) \,dt\biggr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \\
-& \leq \frac{1}{r^{d + 2}}\int_{\Xset} \int_{\Xset} \int_0^1   \|x' - x\|^2 \biggl\{\sum_{i = 1}^{d} \Bigl(f^{(e_i)}\bigl(x + t(x' - x)\bigr)\Bigr)^2\biggr\} K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x)   \,dt   \,dx' \,dx \\
-& \leq \frac{(1 + L_pr)^2}{r^{d + 2}}\int_{\Xset} \int_{\Xset} \int_0^1  \|x' - x\|^2 \biggl\{\sum_{i = 1}^{d} \Bigl(f^{(e_i)}\bigl(x + t(x' - x)\bigr)\Bigr)^2\biggr\} K\biggl(\frac{\|x' - x\|}{r}\biggr) p\bigl(x + t(x' - x)\bigr)^2   \,dt   \,dx' \,dx.
-\end{align*} 
-Then letting $z = (x' - x)/r$ and $\wt{x} = x + trz$, we obtain
-\begin{align*}
-E_r(f) & \leq (1 + L_pr)^2 \int_{\Xset} \int_{\mathcal{Z}_r(x)} \int_{0}^{1} \|z\|^2 \biggl\{\sum_{i = 1}^{d} \Bigl(f^{(e_i)}\bigl(x + trz\bigr)\Bigr)^2\biggr\} K\bigl(\|z\|\bigr) p(x + trz)^2 \,dt \,dz \,dx \\
-& = (1 + L_pr)^2 \int_{B(0,1)} \int_{0}^{1}  \int_{\Xset_r(z)} \|z\|^2 \biggl\{\sum_{i = 1}^{d} \Bigl(f^{(e_i)}\bigl(x + trz\bigr)\Bigr)^2\biggr\} K\bigl(\|z\|\bigr) p(x + trz)^2 \,dx \,dt \,dz \\
-& \leq (1 + L_pr)^2 \int_{B(0,1)} \int_{0}^{1} \|z\|^2 K\bigl(\|z\|\bigr) \biggl\{\int_{\wt{\Xset}_r(t,z)} \|\nabla f(\wt{x})\|^2 p(\wt{x})^2 \,d\wt{x}\biggr\} \,dt \,dz  \\
-& \leq (1 + L_pr)^2 \int_{B(0,1)} \int_{0}^{1} \|z\|^2 K\bigl(\|z\|\bigr) \biggl\{\int_{\Xset} \|\nabla f(\wt{x})\|^2 p(\wt{x})^2 \,d\wt{x}\biggr\} \,dt \,dz \\
-& = (1 + L_pr)^2 \int_{B(0,1)} \|z\|^2 K\bigl(\|z\|\bigr) \,dz 
-\end{align*}
-where $\mathcal{Z}_r(x) = B(0,1) \cap (\Xset - x)/r$, $\Xset_r(z) = \Xset \cap( \Xset - zr)$, $\wt{\Xset}_r(t,z) = \Xset_r(z) + trz$, and we note that since
-\begin{align*}
-\wt{x} \in \wt{\Xset}_r(t,z) & \Longrightarrow \wt{x} = x + trz ~~ \textrm{for some $x \in \Xset - zr$ and $t \in [0,1]$} \\
-& \Longrightarrow \wt{x} \in \Xset,~~ \textrm{(since $x \in \Xset$, $x + rz \in \Xset$ and $\Xset$ is connected)}
-\end{align*}
-it must be that $\wt{X}_r(t,z) \subseteq \Xset$ for all $z \in B(0,1)$ and $t \in [0,1]$.
-
 \paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}.}
-For any $a \in \Reals$, $\Lambda_rf$ satisfies the identity
-\begin{equation*}
-\Lambda_rf(x) = a + \frac{1}{r^d\tau(x)}\int_{\Xset} \eta_r(x',x)\bigl(f(x') - a\bigr)\,dx'
-\end{equation*}
-and by differentiating we obtain
-\begin{equation*}
-\bigl(\nabla \Lambda_rf\bigr)(x)= \frac{1}{r^d\tau(x)}\int_{\Xset} \bigl(\nabla \eta_r(x',\cdot)\bigr)(x)\bigl(f(x') - a\bigr)\,dx' + \nabla\bigl(\tau^{-1}\bigr)(x)\cdot \frac{1}{r^d}\int_{\Xset} \eta_r(x',x)\bigl(f(x') - a\bigr)\,dx'
-\end{equation*} 
-Plugging in $a = f(x)$, we get $\nabla\Lambda_rf(x) = J_1(x)/\tau(x) + J_2(x)$ for
-\begin{equation*}
-J_1(x) := \frac{1}{r^d}\int_{\Xset} \bigl(\nabla \eta_r(x',\cdot)\bigr)(x)\bigl(f(x') - f(x)\bigr)\,dx',~~ J_2(x) := \nabla\bigl(\tau^{-1}\bigr)(x)\cdot \frac{1}{r^d}\int_{\Xset} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)\,dx'.
-\end{equation*}
-To upper bound $\bigl\|J_1(x)\bigr\|^2$, we first compute the gradient of $\eta_r(x',\cdot)$,
-\begin{align*}
-\bigl(\nabla\eta_r(x',\cdot)\bigr)(x) & = \frac{1}{r} \psi'\biggl(\frac{\|x'  - x\|}{r}\biggr) \frac{(x - x')}{\|x' - x\|} \\
-& = \frac{1}{\sigma_Kr^2} K\biggl(\frac{\|x' - x\|}{r}\biggr) (x' - x),
-\end{align*}
-so that by the Cauchy-Schwarz inequality,
-\begin{align*}
-\bigl\|J_1(x)\bigr\|^2 & = \frac{1}{\sigma_K^2 r^{4 + 2d}} \Biggl[\int_{\Xset} \bigl(f(x') - f(x)\bigr)K\biggl(\frac{\|x' - x\|}{r}\biggr)(x' - x)\Biggr]^2 \,dx' \\
-& \leq \frac{1}{\sigma_K^2r^{4 + 2d}} \biggl[\int_{\Xset}K\biggl(\frac{\|x' - x\|}{r}\biggr)\|x' - x\|^2\,dx'\biggr] \biggl[\int_{\Xset}K\biggl(\frac{\|x' - x\|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2\,dx'\biggr] \\
-& \leq \frac{1}{\sigma_K r^{2 + d}}\int_{\Xset}K\biggl(\frac{\|x' - x\|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2\,dx'.
-\end{align*}
-To upper bound $\bigl\|J_2(x)\bigr\|^2$, we use the Cauchy-Schwarz inequality along with the observation $\eta_r(x',x) \leq \frac{1}{\sigma_K} K\bigl(\|x' - x\|/r\bigr)$ to deduce
-\begin{align*}
-\bigl\|J_2(x)\bigr\|^2 & \leq \Bigl\|\nabla\bigl(\tau^{-1}\bigr)(x)\Bigr\|^2\frac{1}{r^{2d}} \biggl[\int_{\Xset}\eta_r(x',x) \,dx'\biggr] \cdot \biggl[\int_{\Xset} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr] \\
-& = \Bigl\|\nabla\bigl(\tau^{-1}\bigr)(x)\Bigr\|^2\frac{\tau(x)}{r^d} \int_{\Xset} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)^2 \,dx'\\ 
-& \leq \Bigl\|\nabla\bigl(\tau^{-1}\bigr)(x)\Bigr\|^2\frac{\tau(x)}{\sigma_K r^d}\int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx' \\
-& \leq \frac{16}{\sigma_K^2r^{2 + d}} \biggl[\int_{\Xset} K\biggl(\frac{|x' - x|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr].
-\end{align*}
-where the last inequality follows by the estimates on $\tau$ and $\nabla \tau$ given in Lemma~\ref{lem:tau_bound}.
-Combining our bounds on $\bigl\|J_1(x)\bigr\|^2$ and $\bigl\|J_2(x)\bigr\|^2$ along with the lower bound on $\tau(x)$ in Lemma~\ref{lem:tau_bound} and integrating over $\Xset$, we have
-\begin{align*}
-\sigma_K D_2(\Lambda_r f) & = \sigma_K\int_{\Xset} \biggl\|\Bigl(\nabla \Lambda_rf)(x)\biggr\|^2 p^2(x) \,dx \\
-& \leq 2 \sigma_K \int_{\Xset} \Biggl(\frac{\bigl\|J_1(x)\bigr\|^2}{\tau^2(x)} + \bigl\|J_2(x)\bigr\|^2\Biggr) p^2(x) \,dx \\
-& \leq \frac{40}{\sigma_K r^{2 + d}} \int_{\Xset} \int_{\Xset} K\biggl(\frac{|x' - x|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 p^2(x) \,dx' \,dx \\
-& \leq \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} E_r(f),
-\end{align*}
-completing the proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}.
+For any $a \in \Reals$, $\Lambda_hf$ satisfies the identity
+\begin{equation*}
+\Lambda_hf(x) = a + \frac{1}{h^d\tau_h(x)}\int_{\Xset} \eta_h(x',x)\bigl(f(x') - a\bigr)\,dx'
+\end{equation*}
+and by differentiating with respect to $x$,  we obtain
+\begin{equation*}
+\bigl(\nabla \Lambda_hf\bigr)(x)= \frac{1}{h^d\tau_h(x)}\int_{\Xset} \bigl(\nabla \eta_h(x',\cdot)\bigr)(x)\bigl(f(x') - a\bigr)\,dx' + \nabla\biggl(\frac{1}{\tau_h}\biggr)(x)\cdot \frac{1}{h^d}\int_{\Xset} \eta_h(x',x)\bigl(f(x') - a\bigr)\,dx'
+\end{equation*} 
+Plugging in $a = f(x)$, we get $\nabla\Lambda_hf(x) = J_1(x)/\tau_h(x) + J_2(x)$ for
+\begin{equation*}
+J_1(x) := \frac{1}{h^d}\int_{\Xset} \bigl(\nabla \eta_h(x',\cdot)\bigr)(x)\bigl(f(x') - f(x)\bigr)\,dx',~~ J_2(x) := \nabla\biggl(\frac{1}{\tau_h}\biggr)(x)\cdot \frac{1}{h^d}\int_{\Xset} \eta_h(x',x)\bigl(f(x') - f(x)\bigr)\,dx'.
+\end{equation*}
+To upper bound $\bigl\|J_1(x)\bigr\|^2$, we first compute the gradient of $\eta_h(x',\cdot)$,
+\begin{align*}
+\bigl(\nabla\eta_h(x',\cdot)\bigr)(x) & = \frac{1}{h^{d + 1}} \psi'\biggl(\frac{\|x'  - x\|}{h}\biggr) \frac{(x - x')}{\|x' - x\|} \\
+& = \frac{1}{\sigma_Kh^{d + 2}} K\biggl(\frac{\|x' - x\|}{h}\biggr) (x' - x),
+\end{align*}
+and additionally note that $\|J_1(x)\|^2 = \sup_{w}\bigl(\langle J_1(x), w \rangle\bigr)^2$ where the supremum is over unit norm vector.
+Taking $w$ to be a unit norm vector which achieves this supremum, we have that
+\begin{align*}
+\bigl\|J_1(x)\bigr\|^2 & = \frac{1}{\sigma_K^2 h^{4 + 2d}} \Biggl[\int_{\Xset} \bigl(f(x') - f(x)\bigr)K\biggl(\frac{\|x' - x\|}{h}\biggr)(x' - x)^Tw\,dx'\Biggr]^2 \\
+& \leq \frac{1}{\sigma_K^2 h^{4 + 2d}} \biggl[\int_{\Xset}K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl((x' - x)^T w\bigr)^2\,dx'\biggr] \biggl[\int_{\Xset}K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl(f(x') - f(x)\bigr)^2\,dx'\biggr].
+\end{align*}
+By a change of variables, we obtain
+\begin{align*}
+\int_{\Xset}K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl((x' - x)^T w\bigr)^2\,dx' & \leq \int_{\Rd}K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl((x' - x)^T w\bigr)^2\,dx' \\
+& = h^{d + 2} \int_{\Rd} K\bigl(\|z\|\bigr) \bigl(z^T w\bigr)^2 \,dz \\
+& = \sigma_K h^{d + 2},
+\end{align*}
+with the resulting upper bound
+\begin{equation*}
+\bigl\|J_1(x)\bigr\|^2 \leq \frac{1}{\sigma_K h^{2 + d}} \int_{\Xset}K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl(f(x') - f(x)\bigr)^2\,dx'.
+\end{equation*}
+To upper bound $\bigl\|J_2(x)\bigr\|^2$, we use the Cauchy-Schwarz inequality along with the observation $\eta_h(x',x) \leq \frac{1}{\sigma_K} K\bigl(\|x' - x\|/h\bigr)$ to deduce
+\begin{align*}
+\bigl\|J_2(x)\bigr\|^2 & \leq \Bigl\|\nabla\biggl(\frac{1}{\tau_h}\biggr)(x)\Bigr\|^2\frac{1}{h^{2d}} \biggl[\int_{\Xset}\eta_h(x',x) \,dx'\biggr] \cdot \biggl[\int_{\Xset} \eta_h(x',x)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr] \\
+& = \Bigl\|\nabla\biggl(\frac{1}{\tau_h}\biggr)(x)\Bigr\|^2\frac{\tau_h(x)}{h^d} \int_{\Xset} \eta_h(x',x)\bigl(f(x') - f(x)\bigr)^2 \,dx'\\ 
+& \leq \Bigl\|\nabla\biggl(\frac{1}{\tau_h}\biggr)(x)\Bigr\|^2\frac{\tau_h(x)}{\sigma_K h^d}\int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx' \\
+& \leq \frac{1}{da_3^3\sigma_K^2h^{2 + d}} \biggl[\int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr].
+\end{align*}
+where the last inequality follows from the estimates on $\tau_h$ and $\nabla \tau_h$ provided in Lemma~\ref{lem:tau_bound}.
+Combining our bounds on $\bigl\|J_1(x)\bigr\|^2$ and $\bigl\|J_2(x)\bigr\|^2$ along with the lower bound on $\tau_h(x)$ in Lemma~\ref{lem:tau_bound} and integrating over $\Xset$, we have
+\begin{align*}
+\sigma_K D_2(\Lambda_h f) & = \sigma_K\int_{\Xset} \biggl\|\Bigl(\nabla \Lambda_hf)(x)\biggr\|^2 p^2(x) \,dx \\
+& \leq 2 \sigma_K \int_{\Xset} \Biggl(\frac{\bigl\|J_1(x)\bigr\|^2}{\tau_h^2(x)} + \bigl\|J_2(x)\bigr\|^2\Biggr) p^2(x) \,dx \\
+& \leq \biggl(\frac{1}{a_3^2} + \frac{1}{da_3^3\sigma_K}\biggr)\frac{2}{h^{d + 2}} \int_{\Xset} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl(f(x') - f(x)\bigr)^2 p^2(x) \,dx' \,dx \\
+& \leq 2 \Bigl(1 + \frac{L_ph}{p_{\min}}\Bigr) \biggl(\frac{1}{a_3^2} + \frac{1}{da_3^3\sigma_K}\biggr) E_h(f),
+\end{align*}
+and taking $A_8 := 2 \Bigl(1 + \frac{L_pc_0}{p_{\min}}\Bigr) \biggl(\frac{1}{a_3^2} + \frac{1}{da_3^3\sigma_K}\biggr)$ completes the proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}.
 
 \paragraph{Proof of Lemma~\ref{lem:tau_bound}.}
-First prove the estimates of $\tau(x)$, and then upper bound on $\|\nabla\tau(x)\|$. Substituting $z = (x' - x)/r$ and using~\ref{asmp:r_small_3}, we have that
-\begin{align*}
-\tau(x) & = \frac{1}{r^d} \int_{\Xset} \psi\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx' \\
-& =  \int_{\mc{B}_x} \psi\bigl(\|z\|\bigr) \,dz \\
-& \geq a_1(\Xset,d) \cdot \int_{B(0,1)} \psi(\|z\|) \,dz
-\end{align*}
-We will now show that $\int_{B(0,1)} \psi(\|z\|) \,dz = 1$, which will imply our estimate of $\tau(x)$.
-To see this identity, note that on the one hand, by converting to polar coordinates and integrating by parts we obtain
-\begin{align*}
-\int_{B(0,1)} \psi\bigl(\|z\|\bigr) \,dz = d \nu_d \int_{0}^{1} \psi(t) t^{d - 1} \,dt = \nu_d \int_{0}^{1} \psi'(t) t^{d} \,dt = \frac{\nu_d}{\sigma_K} \int_{0}^{1} t^{d + 1} K(t) \,dt;
-\end{align*}
-on the other hand, again converting to polar coordinate, we have
-\begin{equation*}
-\sigma_K = \frac{1}{d} \int_{\Reals^d} \|x\|^2 K(\|x\|) \,dx = \nu_d \int_{0}^{1}t^{d + 1} K(t) \,dt.
-\end{equation*}
-
-Finally, we upper bound $\|\nabla\tau(x)\|^2$. Exchanging derivative and integral, we have
-\begin{align*}
-\nabla\tau(x) = \frac{1}{r^d} \int_{\Xset} \bigl(\nabla \eta_r(x',\cdot)\bigr)(x) \,dx' = \frac{1}{\sigma_K r^{d + 2}} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr)(x' - x)\,dx',
-\end{align*}
-and by the Cauchy-Schwarz inequality,
-\begin{equation*}
-\|\nabla\tau(x)\|^2 \leq \frac{1}{\sigma_K^2 r^{2d + 4}} \biggl[\int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr)\,dx'\biggr] \int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr)\|x' - x\|^2\,dx', \leq \frac{1}{\sigma_K r^{2}}
-\end{equation*}
-concluding the proof of Lemma~\ref{lem:tau_bound}. One note: while $\nabla\tau(x) = 0$ when $B(x,r) \in \Xset$, near the boundary the upper bound we derived by using Cauchy-Schwarz appears tight.
+We first establish our estimates of $\tau_h(x)$, and then upper bound $\|\nabla\tau_h(x)\|$. Using~\eqref{eqn:integral_boundary}, we have that
+\begin{align*}
+\tau_h(x) & = \frac{1}{h^d} \int_{\Xset \cap B(x,h)} \psi\biggl(\frac{\|x' - x\|}{h}\biggr) \,dx' \\
+& \geq \frac{a_3}{h^d} \int_{B(x,h)} \psi\biggl(\frac{\|x' - x\|}{h}\biggr) \,dx' \\
+& = a_3\int_{B(0,1)} \psi(\|z\|) \,dz,
+\end{align*}
+and it follows from similar reasoning that $\tau_h(x) \leq \int_{B(0,1)} \psi(\|z\|) \,dz$.
+
+We will now show that $\int_{B(0,1)} \psi(\|z\|) \,dz = 1$, from which we derive the estimates $a_3 \leq \tau_h(x) \leq 1$.
+To see the identity, note that on the one hand, by converting to polar coordinates and integrating by parts we obtain
+\begin{align*}
+\int_{B(0,1)} \psi\bigl(\|z\|\bigr) \,dz = d \nu_d \int_{0}^{1} \psi(t) t^{d - 1} \,dt = -\nu_d \int_{0}^{1} \psi'(t) t^{d} \,dt = \frac{\nu_d}{\sigma_K} \int_{0}^{1} t^{d + 1} K(t) \,dt;
+\end{align*}
+on the other hand, again converting to polar coordinates, we have
+\begin{equation*}
+\sigma_K = \frac{1}{d} \int_{\Reals^d} \|x\|^2 K(\|x\|) \,dx = \nu_d \int_{0}^{1}t^{d + 1} K(t) \,dt,
+\end{equation*}
+and so $\int_{B(0,1)} \psi(\|z\|) \,dz = 1$.
+
+Now we upper bound $\|\nabla\tau_h(x)\|^2$. Exchanging derivative and integral, we have
+\begin{align*}
+\nabla\tau_h(x) = \frac{1}{h^d} \int_{\Xset} \bigl(\nabla \eta_h(x',\cdot)\bigr)(x) \,dx' = \frac{1}{\sigma_K h^{d + 2}} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr)(x' - x)\,dx',
+\end{align*}
+whence by the Cauchy-Schwarz inequality,
+\begin{equation*}
+\|\nabla\tau_h(x)\|^2 \leq \frac{1}{\sigma_K^2 h^{2d + 4}} \biggl[\int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr)\,dx'\biggr] \biggl[\int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr)\|x' - x\|^2\,dx',\biggr] \leq \frac{1}{d\sigma_K h^{2}}
+\end{equation*}
+concluding the proof of Lemma~\ref{lem:tau_bound}.
+
+We remark that while $\nabla\tau(x) = 0$ when $B(x,r) \in \Xset$, near the boundary the upper bound we derived by using Cauchy-Schwarz appears tight.
 
@@
 \begin{align*}
-\Bigl[\Lambda_rf(x)\Bigr]^2 & \leq \frac{1}{r^d \tau(x)}\int_{\Xset} \eta_r(x',x) \bigl[f(x')\bigr]^2 \,dx' \\
-& \leq \frac{2}{r^d p_{\min}}\int_{\Xset} \eta_r(x',x) \bigl[f(x')\bigr]^2 p(x') \,dx'
-\end{align*}
-Then integrating over $x$ and swapping the order of integraton, we have
-\begin{align*}
-\bigl\|\Lambda_rf\bigr\|_{P}^2 & \leq \frac{2}{r^d p_{\min}} \int_{\Xset} \int_{\Xset} \eta_r(x',x) \bigl[f(x')\bigr]^2 p(x') p(x) \,dx' \,dx \\ 
-& \leq \frac{2p_{\max}}{p_{\min}} \int_{\Xset} \bigl[f(x')\bigr]^2 p(x') \,dx' \\
-& = \frac{2p_{\max}}{p_{\min}} \|f\|_{P}^2.
-\end{align*}
-
-Noting that $\Lambda_ra = a$ for any $a \in \Reals$, by the Cauchy-Schwarz inequality we have that for all $x \in \Xset$,
-\begin{align*}
-\bigl|\Lambda_rf(x) - f(x)\bigr|^2 & = \biggl[\frac{1}{r^d\tau(x)} \int_{\Xset} \eta_r(x',x) \bigl(f(x') - f(x)\bigr) \,dx'\biggr]^2 \\
-& \leq \frac{1}{r^{2d} \tau^2(x)} \biggl[\int_{\Xset} \eta_r(x',x) \,dx'\biggr] \cdot \biggl[\int_{\Xset} \eta_r(x',x) \bigl(f(x') - f(x)\bigr)^2 \,dx'\biggr] \\
-& = \frac{1}{r^d \tau(x)} \int_{\Xset} \eta_r(x',x) \bigl(f(x') - f(x)\bigr)^2 \,dx'. \\
-& \leq \frac{1}{r^d \tau(x) p_{\min}} \int_{\Xset} \eta_r(x',x) \bigl(f(x') - f(x)\bigr)^2 p(x') \,dx'.
-\end{align*}
+\Bigl|\Lambda_hf(x)\Bigr|^2 & \leq \frac{1}{h^d\tau_h(x)}\int_{\Xset} \eta_h(x',x) \bigl[f(x')\bigr]^2 \,dx' \\
+& \leq \frac{1}{a_3 h^dp_{\min}}\int_{\Xset} \eta_h(x',x) \bigl[f(x')\bigr]^2 p(x') \,dx'.
+\end{align*}
+Then, integrating over $x$, and recalling that$\int_{B(0,1)} \psi(\|z\|) = 1$ as shown in the proof of Lemma~\ref{lem:tau_bound}, we have
+\begin{align*}
+\bigl\|\Lambda_hf\bigr\|_{P}^2 & \leq \frac{1}{a_3h^d p_{\min}} \int_{\Xset} \int_{\Xset} \eta_h(x',x) \bigl[f(x')\bigr]^2 p(x') p(x) \,dx' \,dx \\ 
+& \leq \frac{p_{\max}}{a_3h^dp_{\min}} \int_{\Xset} \bigl[f(x')\bigr]^2 p(x') \biggl(\int_{\Xset} \eta_h(x',x) \,dx\biggr) \,dx' \\
+& \leq \frac{p_{\max}}{a_3p_{\min}} \int_{\Xset} \bigl[f(x')\bigr]^2 p(x') \biggl(\int_{B(0,1)} \psi(\|z\|) \,dz\biggr) \,dx' \\
+& = \frac{p_{\max}}{a_3p_{\min}} \|f\|_{P}^2.
+\end{align*}
+
+To establish~\eqref{eqn:smoothening_error_energy}, noting that $\Lambda_ha = a$ for any $a \in \Reals$, we have that
+\begin{align*}
+\bigl|\Lambda_rf(x) - f(x)\bigr|^2 & = \biggl[\frac{1}{h^d\tau_h(x)} \int_{\Xset} \eta_h(x',x) \bigl(f(x') - f(x)\bigr) \,dx'\biggr]^2 \\
+& \leq \frac{1}{h^{2d} \tau_h^2(x)} \biggl[\int_{\Xset} \eta_h(x',x) \,dx'\biggr] \cdot \biggl[\int_{\Xset} \eta_h(x',x) \bigl(f(x') - f(x)\bigr)^2 \,dx'\biggr] \\
+& = \frac{1}{h^d \tau_h(x)} \int_{\Xset} \eta_h(x',x) \bigl(f(x') - f(x)\bigr)^2 \,dx'. \\
+& \leq \frac{1}{h^d \tau_h(x) p_{\min}} \int_{\Xset} \eta_h(x',x) \bigl(f(x') - f(x)\bigr)^2 p(x') \,dx'.
+\end{align*}
+From here, we can use the lower bound $\tau_h(x) \geq a_3$ stated in Lemma~\ref{lem:tau_bound}, as well as the upper bound $\eta_h(x',x) \leq (1/\sigma_K) K(\|x' - x\|/h)$, to deduce
+\begin{equation*}
+\bigl|\Lambda_rf(x) - f(x)\bigr|^2 \leq \frac{1}{h^{d} a_3 \sigma_K p_{\min}} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr) \bigl(f(x') - f(x)\bigr)^2 p(x') \,dx'
+\end{equation*}
 Then integrating over $\Xset$ with respect to $p$ yields~\eqref{eqn:smoothening_error_energy}.
@@
 \label{subsec:random_functionals}
-
-We begin by relating $\wt{E}_r(f)$ and $E_r(f)$.
-Note that by assumption $A_2(\wt{\delta} + \theta)/p_{\min} \leq A_1(\wt{\delta} + \theta) \leq \frac{1}{2}$.
-Therefore, some standard calculations show that,
+We will use Lemma~\ref{lem:poincare} in the proof of Proposition~\ref{prop:isometry}. 
+\begin{lemma}[c.f. Lemma 3.4 of \citep{burago2014}]
+	\label{lem:poincare}
+	Let $U \subset \Xset$ be a measurable subset such that $\mathrm{vol}(U) > 0$, and $\diam(U) \leq 2A_0\wt{\delta}$.
+Then, letting $a = (\wt{P}_n(U))^{-1} \cdot \int_{U} f(x) \wt{p}_n(x) \,dx$ be the average of $f$ over $U$, it holds that
+	\begin{equation*}
+	\int_{U} \Bigl|f(x)-a\Bigr|^2 \wt{p}_n(x)\,dx \leq A_3 r^2 \wt{E}_r(f,U).
+	\end{equation*}
+\end{lemma}
+
+Now we relate $\wt{E}_r(f)$ and $E_r(f)$. Some standard calculations show that for $A_1 := 3A_0/p_{\min}$,
 \begin{equation}
@@
 \label{eqn:calder19_2}
-\bigl(1 - A_1(\theta + \wt{\delta})\bigr) \|f\|_{P}^2 \leq \|f\|_{\Leb^2(\wt{P}_n)}^2 \leq \bigl(1 + A_1(\theta + \wt{\delta})\bigr) \|f\|_{P}^2.
+\bigl(1 - A_1(\theta + \wt{\delta})\bigr) \|f\|_{P}^2 \leq \|f\|_{\wt{P}_n}^2 \leq \bigl(1 + A_1(\theta + \wt{\delta})\bigr) \|f\|_{P}^2.
 \end{equation}
@@
 Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized} relates the graph Sobolev semi-norm $b_r(\wt{\mc{P}}f)$ to the non-local energy $\wt{E}_r(f)$. 
-It follows exactly from the derivations in the proof of Lemma 13 in \cite{trillos2019}; indeed, in our flat Euclidean setting, the proof is simplified, and we include this simpler proof for completeness purposes only.
 \begin{lemma}[\textbf{c.f. Lemma 13 of \cite{trillos2019}, Lemma 4.3 of \cite{burago2014}}]
 	\label{lem:first_order_graph_sobolev_seminorm_discretized}
-	Suppose~\ref{asmp:kernel} is satisfied, and that the conclusions of Proposition~\ref{prop:optimal_transport} are met. Then
-	\begin{equation*}
-	b_r(\wt{\mc{P}}f) \leq \wt{E}_{r + 2\wt{\delta}}(f) + 2\frac{L_K\wt{\delta}}{K(1/2)r} \wt{E}_{2(r + 2\wt{\delta})}(f)
-	\end{equation*}
-	for any $f \in \Leb^2(\Xset)$.
+	For any $f \in \Leb^2(\Xset)$,
+	\begin{equation*}
+	b_r(\wt{\mc{P}}f) \leq \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr) \wt{E}_{r + 2A_0\wt{\delta}}(f)
+	\end{equation*}
 \end{lemma}
@@
 	\label{lem:first_order_graph_sobolev_seminorm_discretized_lb}
-	Suppose $\frac{\wt{\delta}}{r} \leq \min\bigl\{\frac{1}{4(2^{d + 2} - 1)}, \frac{K(1)}{2L_K}\bigr\}$. Then for any $u \in \Leb^2(P_n)$, 
-	\begin{equation*}
-	\wt{E}_{r - 2\wt{\delta}}\bigl(\wt{\mc{P}}^{\star}u\bigr) \leq \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) b_{r}(u)
+	For any $u \in \Leb^2(P_n)$, 
+	\begin{equation*}
+	\wt{E}_{r - 2A_0\wt{\delta}}\bigl(\wt{\mc{P}}^{\star}u\bigr) \leq \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) b_{r}(u)
 	\end{equation*}
@@
 
+\paragraph{Proof of Lemma~\ref{lem:poincare}.}
+A symmetrization argument implies that
+\begin{equation}
+\label{pf:poincare_1}
+\int_{U} \Bigl|f(x)-a\Bigr|^2 \wt{p}_n(x)\,dx = \frac{1}{2\wt{P}_n(U)} \int_{U} \int_{U} \bigl|f(x') - f(x)\bigr|^2 \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx
+\end{equation}
+Now, since $x'$ and $x$ belong to $U$, we have that $\|x' - x\| \leq 2A_0\wt{\delta}$.
+Set $V = B(x,r) \cap B(x',r)$, and note that $B(x,r - 2A_0\wt{\delta}) \subset V$.
+Moreover, $r - 2A_0\wt{\delta} \leq r \leq c_0$ by assumption.
+Therefore by~\eqref{eqn:integral_boundary},
+\begin{equation*}
+\vol\bigl(V \cap \Xset\bigr) \geq \vol\bigl(B(x,r - 2A_0\wt{\delta})\bigr) \geq a_3 \nu_d (r - 2A_0\wt{\delta})^d \geq \frac{a_3 \nu_d}{2^d}r^d 
+\end{equation*}
+where the last inequality follows since $\wt{\delta} \leq \frac{1}{4A_0}r$. Using the triangle inequality 
+\begin{equation*}
+\bigl|f(x') - f(x)\big|^2 \leq 2\bigl(\bigl|f(x') - f(z)\big|^2 + \bigl|f(z) - f(x)\big|^2\bigr)
+\end{equation*}
+we have that for any $x$ and $x'$ in $U$,
+\begin{align*}
+\bigl|f(x') - f(x)\big|^2 & \leq \frac{2}{\vol(V \cap \Xset)} \int_{V \cap \Xset} \bigl|f(x') - f(z)\big|^2 + \bigl|f(z) - f(x)\big|^2 \,dz \\
+& \leq \frac{2^{d + 1}}{a_3 \nu_d r^d} \int_{V \cap \Xset} \bigl|f(x') - f(z)\big|^2 + \bigl|f(z) - f(x)\big|^2 \,dz \\
+& \leq \frac{2^{d + 2}}{K(1) a_3 \nu_d r^d p_{\min}} \Bigl(F(x') + F(x)\bigr)
+\end{align*}
+where
+\begin{equation*}
+F(x) := \int_{\Xset} K\biggl(\frac{\|z - x\|}{r}\biggr) \bigl(f(z) - f(x)\bigr)^2 \wt{p}_n(x) \,dx.
+\end{equation*}
+Plugging this upper bound back into~\eqref{pf:poincare_1}, we have that
+\begin{align*}
+\int_{U} \Bigl|f(x)-a\Bigr|^2 \wt{p}_n(x)\,dx & \leq \frac{2^{d + 2}}{K(1) a_3 \nu_d r^d} \int_{U} F(x)\wt{p}_n(x) \,dx \\
+& = \frac{2^{d + 2}}{K(1) a_3 \nu_d}r^2 \wt{E}_r(f,U),
+\end{align*}
+and Lemma~\ref{lem:poincare} follows by taking $A_3 := 2^{d + 2}/(K(1) a_3 \nu_d)$.
+
 \paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}.}
-Recalling that $\bigl(\wt{P}f\bigr)(X_i) = n \cdot \int_{U_i} f(x) \wt{p}_n(x)$, the key point is that
-\begin{equation*}
-\biggl(\bigl(\wt{P}f\bigr)(X_i) - \bigl(\wt{P}f\bigr)(X_j)\biggr)^2 \leq n^2 \cdot \int_{U_i} \int_{U_j} \bigl(f(x') - f(x)\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx.
+Recalling that $\bigl(\wt{\mc{P}}f\bigr)(X_i) = n \cdot \int_{U_i} f(x) \wt{p}_n(x)$, by Jensen's inequality,
+\begin{equation*}
+\biggl(\bigl(\wt{\mc{P}}f\bigr)(X_i) - \bigl(\wt{\mc{P}}f\bigr)(X_j)\biggr)^2 \leq n^2 \cdot \int_{U_i} \int_{U_j} \bigl(f(x') - f(x)\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx.
 \end{equation*}
@@
 \begin{equation*}
-K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) \leq K\biggl(\frac{\bigl(\|x' - x\| - 2\wt{\delta}\bigr)_{+}}{r}\biggr) \leq K\biggl(\frac{\|x' - x\|}{r + 2\wt{\delta}}\biggr) + \frac{2L_K\wt{\delta}}{r}\1\Bigl\{\|x' - x\| \leq r + 2\wt{\delta}\Bigr\}
+K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) \leq K\biggl(\frac{\bigl(\|x' - x\| - 2A_0\wt{\delta}\bigr)_{+}}{r}\biggr) \leq K\biggl(\frac{\|x' - x\|}{r + 2A_0\wt{\delta}}\biggr) + \frac{2L_KA_0\wt{\delta}}{r}\1\Bigl\{\|x' - x\| \leq r + 2A_0\wt{\delta}\Bigr\}
 \end{equation*}
@@
 \begin{align*}
-b_r(f) & = \frac{1}{n^2r^{d + 2}} \sum_{i,j = 1}^n \Bigl(\bigl(\wt{P}f\bigr)(X_i) - \bigl(\wt{P}f\bigr)(X_j)\Bigr)^2 K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) \\
+b_r(\wt{\mc{P}}f) & = \frac{1}{n^2r^{d + 2}} \sum_{i,j = 1}^n \Bigl(\bigl(\wt{\mc{P}}f\bigr)(X_i) - \bigl(\wt{P}f\bigr)(X_j)\Bigr)^2 K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) \\
 & \leq \frac{1}{r^{d + 2}} \sum_{i,j = 1}^n \int_{U_i} \int_{U_j}  \bigl(f(x') - f(x)\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) \,dx' \,dx \\
-& \leq \frac{1}{r^{d + 2}} \sum_{i,j = 1}^n \int_{U_i} \int_{U_j}  \bigl(f(x') - f(x)\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) \biggl[K\biggl(\frac{\|x' - x\|}{r + 2\wt{\delta}}\biggr) + \frac{2L_K\wt{\delta}}{r}\1\Bigl\{\|x' - x\| \leq r + 2\wt{\delta}\Bigr\}\biggr] \,dx' \,dx \\
-& = \wt{E}_{r + 2\wt{\delta}}(f) + \frac{2L_K\wt{\delta}}{r}\wt{E}_{r + 2\wt{\delta}}(f; \1_{[0,1]})
-\end{align*}
-for $\1_{[0,1]}(t) = \1\{0 \leq t \leq 1\}$. But clearly $\wt{E}_{r + 2\wt\delta}(f; \1_{[0,1]}) \leq 1/(K(1/2))\wt{E}_{2r + 4\wt{\delta}}(f)$, and the Lemma is shown.
+& \leq \frac{1}{r^{d + 2}} \sum_{i,j = 1}^n \int_{U_i} \int_{U_j}  \bigl(f(x') - f(x)\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) \biggl[K\biggl(\frac{\|x' - x\|}{r + 2A_0\wt{\delta}}\biggr) + \frac{2L_KA_0\wt{\delta}}{r}\1\Bigl\{\|x' - x\| \leq r + 2\wt{\delta}\Bigr\}\biggr] \,dx' \,dx \\
+& = \Bigl(1 + 2A_0\frac{\wt{\delta}}{r}\Bigr)^{d + 2}\biggl[\wt{E}_{r + 2A_0\wt{\delta}}(f) + \frac{2L_KA_0\wt{\delta}}{r}\wt{E}_{r + 2A_0\wt{\delta}}(f; \1_{[0,1]})\biggr]
+\end{align*}
+for $\1_{[0,1]}(t) = \1\{0 \leq t \leq 1\}$. But by assumption $\wt{E}_{r + 2A_0\wt\delta}(f; \1_{[0,1]}) \leq 1/(K(1))\wt{E}_{r + 2A_0\wt{\delta}}(f)$, and so we obtain
+\begin{equation*}
+b_r(\wt{\mc{P}}f) \leq \Bigl(1 + 2A_0\frac{\wt{\delta}}{r}\Bigr)^{d + 2} \Bigl(1 + \frac{2L_KA_0\wt{\delta}}{rK(1)}\Bigr) \wt{E}_{r + 2A_0\wt{\delta}}(f);
+\end{equation*}
+the Lemma follows upon choosing $A_9 := A_0(2^{d + 4} + \frac{4L_K}{K(1)})$.
 
 \paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}.}
-For brevity, we write $\wt{r} := r - 2\wt{\delta}$. We begin by expanding the energy $\wt{E}_{r - 2\wt{\delta}}\bigl(\wt{\mc{P}}^{\star}u\bigr)$ as a double sum of double integrals,
+For brevity, we write $\wt{r} := r - 2A_0\wt{\delta}$. We begin by expanding the energy $\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr)$ as a double sum of double integrals,
 \begin{align*}
@@
 \begin{equation*}
-K\biggl(\frac{\|x' - x\|}{\wt{r}}\biggr) \leq K\biggl(\frac{\|x_i - x_j\|}{r}\biggr) + \frac{2L_K\wt{\delta}}{\wt{r}} \cdot \1\biggl\{\frac{\|x' - x\|}{\wt{r}} \leq 1\biggr\}
+K\biggl(\frac{\|x' - x\|}{\wt{r}}\biggr) \leq K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) + \frac{2A_0L_K\wt{\delta}}{\wt{r}} \cdot \1\biggl\{\frac{\|x' - x\|}{\wt{r}} \leq 1\biggr\}
 \end{equation*}
@@
 \begin{align}
-\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & \leq \frac{1}{n^2\wt{r}^{d + 2}} \sum_{i = 1}^{n} \sum_{j = 1}^{n} \Bigl(u(X_i) - u(X_j)\Bigr)^2 K\biggl(\frac{\|x_i - x_j\|}{r}\biggr) + \frac{2L_K\wt{\delta}}{\wt{r}}\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \nonumber \\
-& \leq \biggl(1 + 4\bigl(2^{d + 2} - 1\bigr)\frac{\wt{\delta}}{r}\biggr) b_r(u) + \frac{2L_K\wt{\delta}}{\wt{r}}\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \nonumber \\
-& \leq \biggl(1 + 4\bigl(2^{d + 2} - 1\bigr)\frac{\wt{\delta}}{r}\biggr) b_r(u) + \frac{2L_K\wt{\delta}}{K(1)\wt{r}} \wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u\bigr) \nonumber 
+\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & \leq \frac{1}{n^2\wt{r}^{d + 2}} \sum_{i = 1}^{n} \sum_{j = 1}^{n} \Bigl(u(X_i) - u(X_j)\Bigr)^2 K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) + \frac{2A_0L_K\wt{\delta}}{\wt{r}}\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \nonumber \\
+& \leq \biggl(1 + 2^{d + 2}A_0\frac{\wt{\delta}}{r}\biggr) b_r(u) + \frac{2A_0L_K\wt{\delta}}{\wt{r}}\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \nonumber \\
+& \leq \biggl(1 + 2^{d + 2}A_0\frac{\wt{\delta}}{r}\biggr) b_r(u) + \frac{4A_0L_K\wt{\delta}}{K(1)r} \wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u\bigr) \nonumber 
 \end{align}
@@
 \begin{align*}
-\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & \leq \biggl(1 - \frac{2L_K\wt{\delta}}{K(1)\wt{r}}\biggr)^{-1}\biggl(1 + 4\bigl(2^{d + 2} - 1\bigr)\frac{\wt{\delta}}{r}\biggr) b_r(u) \\
-& \leq \biggl(1 + \frac{2L_K}{K(1)} + 8\bigl(2^{d + 2} - 1\bigr)\biggr) \frac{\wt{\delta}}{r}b_r(u)
+\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & \leq \biggl(1 - \frac{4A_0L_K\wt{\delta}}{K(1)r}\biggr)^{-1}\biggl(1 + 2^{d + 2}A_0\frac{\wt{\delta}}{r}\biggr) b_r(u) \\
+& \leq \biggl(1 + \frac{\wt{\delta}}{r}\Bigl(\frac{8A_0L_K}{K(1)} + 2^{d + 3}\Bigr)\biggr) b_r(u)
 \end{align*}
@@
 \begin{align*}
-\sigma_K D_2(\Lambda_{r - 2\wt{\delta}} \wt{\mc{P}}^{\star}u) & \overset{(i)}{\leq} \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} E_{r - 2\wt{\delta}}(\wt{\mc{P}}^{\star}u) \\
-& \overset{(ii)}{\leq} \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \wt{E}_{r - 2\wt{\delta}}(\wt{\mc{P}}^{\star}u) \\
-& \overset{(iii)}{\leq} \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) b_r(u)
+\sigma_K D_2(\Lambda_{r - 2A_0\wt{\delta}} \wt{\mc{P}}^{\star}u) & \overset{(i)}{\leq} A_8 E_{r - 2A_0\wt{\delta}}(\wt{\mc{P}}^{\star}u) \\
+& \overset{(ii)}{\leq} A_8 \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \wt{E}_{r - 2A_0\wt{\delta}}(\wt{\mc{P}}^{\star}u) \\
+& \overset{(iii)}{\leq} A_8 \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) b_r(u)
 \end{align*}
@@
 \begin{align*}
-b_r(\wt{\mc{P}}f) & \overset{(iv)}{\leq} \wt{E}_{r + 2\wt{\delta}}(f) + \frac{2L_K}{K(1/2)}\frac{\wt{\delta}}{r}\wt{E}_{2(r + 2\wt{\delta})}(f)\\
-& \overset{(v)}{\leq} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl({E}_{r + 2\wt{\delta}}(f) + \frac{2L_K}{K(1/2)}\cdot\frac{\wt{\delta}}{r}{E}_{2(r + 2\wt{\delta})}(f)\Bigr) \\
-& \overset{(vi)}{\leq} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + L_p(2r + 4\wt{\delta})\Bigr) \cdot \Bigl(1 + \frac{2L_K}{K(1/2)}\frac{\wt{\delta}}{r}\Bigr) \cdot \sigma_{\eta} D_2(f)
-\end{align*}
-where $(iv)$ follows from Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}, $(v)$ follows from~\eqref{eqn:calder19_1}, and $(vi)$ follows from Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected}.
-The proposition follows upon taking $A_4 := 2L_K/K(1/2)$.
+b_r(\wt{\mc{P}}f) & \overset{(iv)}{\leq} \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr)\wt{E}_{r + 2A_0\wt{\delta}}(f)\\
+& \overset{(v)}{\leq} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr){E}_{r + 2A_0\wt{\delta}}(f) \\
+& \overset{(vi)}{\leq} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + \frac{L_p(r + 2A_0\wt{\delta})}{p_{\min}}\Bigr)^2 \cdot \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr) \cdot \sigma_{K} D_2(f)
+\end{align*}
+where $(iv)$ follows from Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}, $(v)$ follows from~\eqref{eqn:calder19_1}, and $(vi)$ follows from the proof of Lemma~\ref{lem:graph_sobolev_seminorm}.
 
@@
 \textit{Proof of (1).}
-We begin by upper bounding $\bigl\|\wt{P}f\bigr\|_{n}$. By the Cauchy-Schwarz inequality and the bound on $\|\wt{p}_n - p\|_{\infty}$ in~\eqref{eqn:calder19_1},
+We begin by upper bounding $\bigl\|\wt{P}f\bigr\|_{n}$. By the Cauchy-Schwarz inequality and the bound on $\|\wt{p}_n - p\|_{\infty}$ in~\eqref{eqn:optimal_transport_2},
 \begin{align*}
@@
 & \leq n \int_{U_i} \bigl|f(x)\bigr|^2 \wt{p}_n(x) \,dx \\
-& \leq n \cdot \biggl[\int_{U_i} \bigl|f(x)\bigr|^2 p(x) \,dx + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \int_{U_i} \bigl|f(x)\bigr|^2 p(x) \,dx\biggr]
+& \leq n \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)\biggl[\int_{U_i} \bigl|f(x)\bigr|^2 p(x) \,dx + A_1(\theta + \wt{\delta}) \int_{U_i} \bigl|f(x)\bigr|^2 p(x) \,dx\biggr]
 \end{align*}
@@
 \label{pf:prop_isometry_1}
-\bigl\|\wt{P}f\bigr\|_{n}^2 \leq \biggl(1 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \bigl\|f\bigr\|_{P}^2.
-\end{equation}
-Now, noticing that $\bigl\|\wt{P}f\bigr\|_{n} = \bigl\|\wt{P}^{\star}\wt{P}f\bigr\|_{\Leb^2(\wt{P}_n)}$, we can use the upper bound~\eqref{pf:prop_isometry_1} to show that
+\bigl\|\wt{P}f\bigr\|_{n}^2 \leq \biggl(1 + A_1(\theta + \wt{\delta})\biggr) \bigl\|f\bigr\|_{P}^2.
+\end{equation}
+Now, noticing that $\bigl\|\wt{P}f\bigr\|_{n} = \bigl\|\wt{P}^{\star}\wt{P}f\bigr\|_{\wt{P}_n}$, we can use the upper bound~\eqref{pf:prop_isometry_1} to show that
 \begin{align}
-\Bigl|\bigl\|\wt{P}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{P}^2\Bigr| & \leq \Bigl|\bigl\|\wt{P}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2\Bigr| + \Bigl|\bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2 - \bigl\|f\bigr\|_{P}^2\Bigr| \nonumber \\
-& \overset{(i)}{\leq} \Bigl|\bigl\|\wt{P}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2\Bigr|  + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{P}^2 \\
-& \overset{(ii)}{\leq} \biggl(2 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \Bigl|\bigl\|\wt{P}f\bigr\|_{n} - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}\Bigr| \cdot \bigl\|f\bigr\|_{P} + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{P}^2 \nonumber \\
-& \leq \biggl(2 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)} \cdot \bigl\|f\bigr\|_{P} + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{P}^2 \label{pf:prop_isometry_2}
+\Bigl|\bigl\|\wt{P}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{P}^2\Bigr| & \leq \Bigl|\bigl\|\wt{P}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{\wt{P}_n}^2\Bigr| + \Bigl|\bigl\|f\bigr\|_{\wt{P}_n}^2 - \bigl\|f\bigr\|_{P}^2\Bigr| \nonumber \\
+& \overset{(i)}{\leq} \Bigl|\bigl\|\wt{P}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{\wt{P}_n}^2\Bigr|  + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2 \\
+& \overset{(ii)}{\leq} 2 \sqrt{1 + A_1(\theta + \wt{\delta})} \Bigl|\bigl\|\wt{P}f\bigr\|_{n} - \bigl\|f\bigr\|_{\wt{P}_n}\Bigr| \cdot \bigl\|f\bigr\|_{P} + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2 \nonumber \\
+& \leq 2 \sqrt{1 + A_1(\theta + \wt{\delta})} \bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\wt{P}_n} \cdot \bigl\|f\bigr\|_{P} + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2 \label{pf:prop_isometry_2}
 \end{align}
-where $(i)$ follows from~\eqref{eqn:calder19_2} and $(ii)$ follows from~\eqref{pf:prop_isometry_1}.
-
-It remains to upper bound $\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)}^2$. Noting that $\wt{P}^{\star}\wt{P}f$ is piecewise constant over the cells $U_i$, we have
-\begin{equation*}
-\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)}^2 = \sum_{i = 1}^{n} \int_{U_i} \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx.
-\end{equation*}
-We will now establish the following Poincare-type inequality, 
-\begin{equation}
-\label{pf:prop_isometry_3}
-\int_{U_i} \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx \leq \frac{2^{d + 2}}{a_1(\Xset,d) \nu_d K(1/2)} \wt{\delta}^2 \wt{E}_{2\wt{\delta}}(f,U_i)
-\end{equation}
-which holds for all $i = 1,\ldots,n$. Summing up over $i$ implies
-\begin{equation*}
-\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)}^2 \leq \frac{2^{d+2}}{a_1(\Xset,d)K(1/2)p_{\min}} \cdot \wt{E}_{2\wt{\delta}}(f) \leq \frac{2^{d+3}}{a_1(\Xset,d)K(1/2)p_{\min}} A_1(\Xset,d,K,p_{\min}) D_2(f)
-\end{equation*}
-with the latter inequality following just as in the proof of Proposition~\ref{prop:dirichlet_energies}; the Lemma then follows by plugging this inequality into~\eqref{pf:prop_isometry_2}.
-
-Now it remains to show~\eqref{pf:prop_isometry_3}. The key point is that for any pair of points $x,x' \in \Xset$ such that $\|x - x'\| \leq r$  the triangle inequality
-\begin{equation*}
-\bigl|f(x') - f(x)\bigr|^2 \leq 2\bigl|f(x') - f(z)\bigr|^2 + 2\bigl|f(z) - f(x)\bigr|^2
-\end{equation*}
-holds for each $z \in B\bigl(x,\wt{\delta}\bigr) \cap B\bigl(x',\wt{\delta}\bigr) \cap \Xset =: W$, and we have that
-\begin{align*}
-\bigl|f(x') - f(x)\bigr|^2 & \leq \frac{2}{\vol(W)}\biggl[\int_{W} \bigl|f(x') - f(z)\bigr|^2 \,dz + \int_{W} \bigl|f(z) - f(x)\bigr|^2 \,dz\biggr] \\
-& \leq \frac{2}{\vol(W)K(1/2)}\biggl[\int_{\Xset} K\biggl(\frac{x' - z}{2\wt{\delta}}\biggr)\bigl|f(x') - f(z)\bigr|^2 \,dz + \int_{\Xset} K\biggl(\frac{z - x}{2\wt{\delta}}\biggr)\bigl|f(z) - f(x)\bigr|^2 \,dz\biggr] \\
-& \leq \frac{2^{d + 1}}{a_1(\Xset,d) \nu_d \wt{\delta}^d K(1/2)}\biggl[\int_{\Xset} K\biggl(\frac{x' - z}{2\wt{\delta}}\biggr)\bigl|f(x') - f(z)\bigr|^2 \,dz + \int_{\Xset} K\biggl(\frac{z - x}{2\wt{\delta}}\biggr)\bigl|f(z) - f(x)\bigr|^2 \,dz\biggr]
-\end{align*}
-where the last inequality follows from Assumption~\ref{asmp:r_small_1}. We use this along with the Cauchy-Schwarz inequality to prove~\eqref{pf:prop_isometry_3}:
-\begin{align*}
-\int_{U_i} & \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx \\
-& =\int_{U_i} \biggl(n\cdot\int_{U_i} \bigl(f(x) - f(x')\bigr) \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx \\
-& \leq n \cdot \int_{U_i} \int_{U_i} \bigl(f(x) - f(x')\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx \\
-& \leq \frac{2^{d + 2}}{a_1(\Xset,d) \nu_d \wt{\delta}^d K(1/2)}n\int_{U_i} \int_{U_i} \int_{\Xset} K\biggl(\frac{x - z}{2\wt{\delta}}\biggr)\bigl|f(x) - f(z)\bigr|^2 \wt{\rho}_n(x') \wt{\rho}_n(x)  \,dx' \,dz \,dx \\
-& \leq \frac{2^{d + 2}}{a_1(\Xset,d) \nu_d \wt{\delta}^d K(1/2) p_{\min}} \int_{U_i} \int_{U_i} K\biggl(\frac{x - z}{2\wt{\delta}}\biggr)\bigl|f(x) - f(z)\bigr|^2 \wt{\rho}_n(z) \wt{\rho}_n(x)  \,dz \,dx \\
-& =  \frac{2^{d + 2}}{a_1(\Xset,d) \nu_d K(1/2) p_{\min}} \wt{\delta}^2 \wt{E}_{2\wt{\delta}}(f,U_i).
-\end{align*}
+where $(i)$ follows from~\eqref{eqn:calder19_2} and $(ii)$ follows from~\eqref{eqn:calder19_2} and ~\eqref{pf:prop_isometry_1}.
+
+It remains to upper bound $\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\wt{P}_n}^2$. Noting that $\wt{P}^{\star}\wt{P}f$ is piecewise constant over the cells $U_i$, we have
+\begin{equation*}
+\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\wt{P}_n}^2 = \sum_{i = 1}^{n} \int_{U_i} \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx.
+\end{equation*}
+From Lemma~\ref{lem:poincare}, we have that
+\begin{equation*}
+\int_{U_i} \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx \leq A_3r^2 \wt{E}_r(f,U_i)
+\end{equation*}
+for each $i = 1,\ldots,n$. Summing up over $i$ on both sides of the inequality
+\begin{equation*}
+\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\wt{P}_n}^2 \leq A_3r^2 \wt{E}_r(f,\Xset) \leq  A_3 \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + \frac{L_p(r + 2A_0\wt{\delta})}{p_{\min}}\Bigr)^2r^2 D_2(f)
+\end{equation*}
+where the latter inequality follows from the proof of Proposition~\ref{prop:dirichlet_energies}, Part (2).
+Then Proposition~\ref{prop:isometry}, Part (1) follows by plugging this inequality into~\eqref{pf:prop_isometry_2} and taking
+\begin{equation*}
+A_5 := 2\sqrt{A_3} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \Bigl(1 + \frac{L_p(r + 2A_0\wt{\delta})}{p_{\min}}\Bigr).
+\end{equation*}
 
 \textit{Proof of (2).}
-By the triangle inequality,
+By the triangle inequality and~\eqref{eqn:calder19_2},
 \begin{align}
-\Bigl|\|\wt{\mc{I}}u\|_{P}^2 - \|u\|_{n}^2\Bigr| & \leq \Bigl|\|\wt{\mc{I}}u\|_{P}^2 - \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2\Bigr| + \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 - \|u\|_{n}^2\Bigr| \nonumber \\
-& \leq A_1(\theta + \wt{\delta}) \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 + \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 - \|u\|_{n}^2\Bigr| \nonumber\\
-& = A_1(\theta + \wt{\delta}) \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 + \Bigl(\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} + \|u\|_{n}\Bigr) \cdot \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|u\|_{n}\Bigr| \label{pf:prop_isometry_4}
+\Bigl|\|\wt{\mc{I}}u\|_{P}^2 - \|u\|_{n}^2\Bigr| & \leq \Bigl|\|\wt{\mc{I}}u\|_{P}^2 - \|\wt{\mc{I}}u\|_{\wt{P}_n}^2\Bigr| + \Bigl|\|\wt{\mc{I}}u\|_{\wt{P}_n}^2 - \|u\|_{n}^2\Bigr| \nonumber \\
+& \leq A_1(\theta + \wt{\delta}) \|\wt{\mc{I}}u\|_{\wt{P}_n}^2 + \Bigl|\|\wt{\mc{I}}u\|_{\wt{P}_n}^2 - \|u\|_{n}^2\Bigr| \nonumber\\
+& = A_1(\theta + \wt{\delta}) \|\wt{\mc{I}}u\|_{\wt{P}_n}^2 + \Bigl(\|\wt{\mc{I}}u\|_{\wt{P}_n} + \|u\|_{n}\Bigr) \cdot \Bigl|\|\wt{\mc{I}}u\|_{\wt{P}_n} - \|u\|_{n}\Bigr| \label{pf:prop_isometry_4}
 \end{align}
-To upper bound the second term in the above expression, we first note that~$\|u\|_{n} = \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}$, and thus
+To upper bound the second term in the above expression, we first note that~$\|u\|_{n} = \|\wt{\mc{P}}^{\star}u\|_{\wt{P}_n}$, and thus
 \begin{align}
-\Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|u\|_{n} \Bigr| & = \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}\Bigr| \nonumber \\
-& \overset{(iii)}{\leq} \|\Lambda_{\wt{r}}\wt{\mc{P}}^{\star}u - \wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)} \nonumber \\
-& \overset{(iv)}{\leq} r \sqrt{\frac{2}{\sigma_K p_{\min}} E_{\wt{r}}(\wt{\mc{P}}^{\star}u)} \nonumber \\
-& \overset{(v)}{\leq} r \sqrt{\frac{2}{\sigma_K p_{\min}} \Bigl(1 + A_3\frac{\wt{\delta}}{r}\Bigr) b_r(u)} \label{pf:prop_isometry_5}
+\Bigl|\|\wt{\mc{I}}u\|_{\wt{P}_n} - \|u\|_{n} \Bigr| & = \Bigl|\|\wt{\mc{I}}u\|_{\wt{P}_n} - \|\wt{\mc{P}}^{\star}u\|_{\wt{P}_n}\Bigr| \nonumber \\
+& \overset{(iii)}{\leq} \|\Lambda_{\wt{r}}\wt{\mc{P}}^{\star}u - \wt{\mc{P}}^{\star}u\|_{\wt{P}_n} \nonumber \\
+& \overset{(iv)}{\leq} \wt{r} \sqrt{\frac{1}{a_3\sigma_K p_{\min}} E_{\wt{r}}(\wt{\mc{P}}^{\star}u)} \nonumber \\
+& \overset{(v)}{\leq} \wt{r} \sqrt{\frac{1 + A_1(\theta + \wt{\delta})}{a_3\sigma_K p_{\min}} \Bigl(1 + A_3\frac{\wt{\delta}}{r}\Bigr) b_r(u)} \label{pf:prop_isometry_5}
 \end{align}
@@
 \begin{align*}
-\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 & \leq \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \|\wt{\mc{I}}u\|_{P}^2 \\
-& \leq \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \|\wt{\mc{P}}^{\star}u\|_{P}^2 \\
-& \leq \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)^2 \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}^2 \\
-& = \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)^2 \|u\|_{n}^2
-\end{align*}
-and plugging this estimate along with~\eqref{pf:prop_isometry_5} back into~\eqref{pf:prop_isometry_4}, we obtain part (2) of Proposition~\ref{prop:isometry}.
-
-\subsection{\textcolor{red}{(TODO)}: Move these.}
-
-In the above we used the following:
-\begin{enumerate}[label=(OT\arabic*)]
-	\item
-	\label{asmp:r_small_1} For all $x,x' \in \Xset$ such that $\|x - x'\| \leq r$, we have that for $W = B(x,r) \cap B(x',r) \cap \Xset$,
-	\begin{equation*}
-	\vol(W) \geq a_1(\Xset,d) \cdot \nu_d \biggl(\frac{r}{2}\biggr)^d
-	\end{equation*}
-	for some constant $a_1(\Xset,d)$. 
-	\item 
-	\label{asmp:r_small_3} Set $\mc{B}_x:= (\Xset - x)/r~ \cap~B(0,1)$. For any $x \in \Xset$ and any function $\psi: [0,\infty) \to [0,\infty)$,
-	\begin{equation}
-	\label{eqn:lb_expected_degree}
-	\int_{\mc{B}_x} \psi(\|z\|) \,dz \geq a_1(\Xset,d) \cdot \int_{B(0,1)} \psi(\|z\|) \,dz.
-	\end{equation}
-\end{enumerate}
-
-\section{Bounds on the empirical norm}
+\|\wt{\mc{I}}u\|_{\wt{P}_n}^2 & \leq \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \|\wt{\mc{I}}u\|_{P}^2 \\
+& \leq \frac{p_{\max}}{a_3p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \|\wt{\mc{P}}^{\star}u\|_{P}^2 \\
+& \leq \frac{p_{\max}}{a_3p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)^2 \|\wt{\mc{P}}^{\star}u\|_{\wt{P}_n}^2 \\
+& = \frac{p_{\max}}{a_3p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)^2 \|u\|_{n}^2
+\end{align*}
+and plugging this estimate along with~\eqref{pf:prop_isometry_5} back into~\eqref{pf:prop_isometry_4}, upon choosing
+\begin{equation*}
+A_6 := \biggl(3\sqrt{\frac{2p_{\max}}{p_{\min}}} + 1\biggr)\sqrt{\frac{4}{a_3\sigma_Kp_{\min}}},~~A_7:=4A_1\frac{p_{\max}}{a_3p_{\min}}
+\end{equation*}
+we obtain part (2) of Proposition~\ref{prop:isometry}.
+
+\section{Bound on the empirical norm}
 \label{sec:empirical_norm}
 
-We use Lemma~\ref{lem:empirical_norm_sobolev} to lower bound $\norm{f_0}_n^2$ by (a constant times) the $\Leb^2(\Xset)$ norm of $f$.
+In Lemma~\ref{lem:empirical_norm_sobolev}, we lower bound $\norm{f_0}_n^2$ by (a constant times) the $\Leb^2(\Xset)$ norm of $f$.
 
@@
 \end{equation*}
-We will use a Sobolev inequality to relate $\norm{f}_{\Leb^4}$ to $\norm{f}_{H^1(\Xset)}$.
-The nature of this inequality depends on the relationship between $s$ and $d$ (see Theorem 6 in Section 5.6.3 of \cite{evans10} for a formal statement), so from this point on we divide our analysis into three cases: (i) the case where $2s > d$, (ii) the case where $2s < d$, and (iii) the borderline case $2s = d$.
+We will use a Sobolev inequality to relate $\norm{f}_{\Leb^4(\Xset)}$ to $\norm{f}_{H^1(\Xset)}$.
+The nature of this inequality depends on the relationship between $s$ and $d$ (see Theorem 6 in Section 5.6.3 of \cite{evans10} for a formal statement), so from this point on we divide our analysis into three cases: (i) the case where $2 > d$, (ii) the case where $2 < d$, and (iii) the borderline case $2 = d$.
 
 \textit{Case 1: $2s > d$.}
-When $2s > d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in C^{\gamma}(\overline{\Xset})$ for some $\gamma > 0$ which depends on $s$ and $d$, with the accompanying estimate
-\begin{equation*}
-\sup_{x \in \Xset} \abs{f(x)} \leq \norm{f}_{C^{\gamma}(\Xset)} \leq c \norm{f}_{H^s(\Xset)}.
+When $2 > d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in C^{\gamma}(\overline{\Xset})$ for some $\gamma > 0$ which depends on $d$, with the accompanying estimate
+\begin{equation*}
+\sup_{x \in \Xset} \abs{f(x)} \leq \norm{f}_{C^{\gamma}(\Xset)} \leq c \norm{f}_{H^1(\Xset)}.
 \end{equation*}
@@
 \begin{align*}
-\norm{f}_{\Leb^4}^4 & = \int_{\Xset} [f(x)]^4 \,dx \\
+\norm{f}_{\Leb^4(\Xset)}^4 & = \int_{\Xset} [f(x)]^4 \,dx \\
 & \leq \left(\sup_{x \in \Xset} \abs{f(x)}\right)^2 \cdot \int_{\Xset} [f(x)]^2 \,dx \\
-& \leq c \norm{f}_{H^s(\Xset)}^2 \cdot \norm{f}_{\Leb^2(\Xset)}^2.
+& \leq c \norm{f}_{H^1(\Xset)}^2 \cdot \norm{f}_{\Leb^2(\Xset)}^2.
 \end{align*}
@@
 \begin{equation*}
-\norm{f}_{\Leb^2(\Xset)}^2 \geq c_1^2 \cdot b^2 \cdot \norm{f}_{W_d^{s,2}(\Xset)}^2 \cdot \frac{1}{n},
+\norm{f}_{\Leb^2(\Xset)}^2 \geq c_1^2 \cdot b^2 \cdot \norm{f}_{H^{1}(\Xset)}^2 \cdot \frac{1}{n},
 \end{equation*}
@@
 \begin{equation*}
-p_{\max}^2\frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c \norm{f}_{H^s(\Xset)}^2 \cdot \frac{\norm{f}_{\Leb^2(\Xset)}^4}{n \norm{f}_{\Leb^2(\Xset)}^2} \leq c \frac{\norm{f}_{\Leb^2(\Xset)}^4}{c_1^2 b^2} \leq \frac{\Ebb\bigl[\norm{f}_n^2\bigr]}{b^2},
+p_{\max}^2\frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c \norm{f}_{H^1(\Xset)}^2 \cdot \frac{\norm{f}_{\Leb^2(\Xset)}^4}{n \norm{f}_{\Leb^2(\Xset)}^2} \leq c \frac{\norm{f}_{\Leb^2(\Xset)}^4}{c_1^2 b^2} \leq \frac{\Ebb\bigl[\norm{f}_n^2\bigr]}{b^2},
 \end{equation*}
@@
 
-\textit{Case 2: $2s < d$.}
-When $2s < d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in \Leb^q(\Xset)$ for $q = 2d/(d - 2s)$, and moreover that
-\begin{equation*}
-\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{H^s(\Xset)}.
-\end{equation*}
-Since $4 = 2\theta + (1 - \theta)q$ for $\theta = 2 - d/(2s)$, Lyapunov's inequality implies
-\begin{equation*}
-\norm{f}_{\Leb^4(\Xset)}^4 \leq \norm{f}_{\Leb^2}^{2\theta} \cdot \norm{f}_{\Leb^q(\Xset)}^{(1 - \theta)q} \leq c \norm{f}_{\Leb^2(\Xset)}^{4} \cdot \left(\frac{\norm{f}_{H^s(\Xset)}}{\norm{f}_{\Leb^2(\Xset)}}\right)^{d/s}.
-\end{equation*}
-By assumption, $\norm{f}_{\Leb^2(\Xset)} \geq c_1 b \norm{f}_{H^s(\Xset)} n^{-s/d}$, and therefore
-\begin{equation*}
-p_{\max}^2 \frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c\norm{f}_{\Leb^2(\Xset)}^4 \left(\frac{\norm{f}_{H^s(\Xset)}}{n^{s/d}\norm{f}_{\Leb^2(\Xset)}}\right)^{d/s} \leq \frac{c\norm{f}_{\Leb^2(\Xset)}^4}{c_1b^{d/s}} \leq \frac{\norm{f}_{\Leb^2(\Xset)}^4}{b^2}.
-\end{equation*}
-where the last inequality follows when $c_1$ is sufficiently large, and keeping in mind that $d/s > 2$ and $b \geq 1$.
-
-\textit{Case 3: $2s = d$.}
+\textit{Case 2: $2 < d$.}
+When $2 < d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in \Leb^q(\Xset)$ for $q = 2d/(d - 2)$, and moreover that
+\begin{equation*}
+\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{H^1(\Xset)}.
+\end{equation*}
+Since $4 = 2\theta + (1 - \theta)q$ for $\theta = 2 - d/2$, Lyapunov's inequality implies
+\begin{equation*}
+\norm{f}_{\Leb^4(\Xset)}^4 \leq \norm{f}_{\Leb^2}^{2\theta} \cdot \norm{f}_{\Leb^q(\Xset)}^{(1 - \theta)q} \leq c \norm{f}_{\Leb^2(\Xset)}^{4} \cdot \left(\frac{\norm{f}_{H^1(\Xset)}}{\norm{f}_{\Leb^2(\Xset)}}\right)^{d}.
+\end{equation*}
+By assumption, $\norm{f}_{\Leb^2(\Xset)} \geq c_1 b \norm{f}_{H^1(\Xset)} n^{-1/d}$, and therefore
+\begin{equation*}
+p_{\max}^2 \frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c\norm{f}_{\Leb^2(\Xset)}^4 \left(\frac{\norm{f}_{H^1(\Xset)}}{n^{1/d}\norm{f}_{\Leb^2(\Xset)}}\right)^{d} \leq \frac{c\norm{f}_{\Leb^2(\Xset)}^4}{c_1b^{d}} \leq \frac{\norm{f}_{\Leb^2(\Xset)}^4}{b^2}.
+\end{equation*}
+where the last inequality follows when $c_1$ is sufficiently large, and keeping in mind that $d > 2$ and $b \geq 1$.
+
+\textit{Case 3: $2 = d$.}
 Assume $f$ satisfies~\eqref{eqn:paley_zygmund_1} for a given $0 < a < 1$.
-When $2s = d$, since $\Xset$ is a Lipschitz domain we have that $f \in L^q(\Xset)$ for any $q < \infty$, with the accompanying estimate
-\begin{equation*}
-\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{H^s(\Xset)}.
+When $2 = d$, since $\Xset$ is a Lipschitz domain we have that $f \in L^q(\Xset)$ for any $q < \infty$, with the accompanying estimate
+\begin{equation*}
+\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{H^1(\Xset)}.
 \end{equation*}
@@
 
-\section{Proof of Main Results}
+\section{Graph functionals under the manifold hypothesis}
+\label{sec:manifold}
+In this section, we restate a few results of \citep{trillos2019,calder2019}, which are analogous to Lemmas~\ref{lem:graph_sobolev_seminorm} and~\ref{lem:neighborhood_eigenvalue} but cover the case where $\Xset$ is an $m$-dimensional submanifold without boundary.
+As such, the results in this section will hold under the assumption~\ref{asmp:domain_manifold}.
+We refer to~\citep{trillos2019,calder2019} for the proofs of these results.
+
+Proposition~\ref{prop:garciatrillos19_1} follows from Lemma~5 of~\citep{trillos2019} and Markov's inequality.
+\begin{proposition}[]
+	\label{prop:garciatrillos19_1}
+	For any $f \in H^1(\Xset)$, with probability at least $1 - \delta$,
+	\begin{equation*}
+	f^T \Lap_{n,r} f \leq \frac{C}{\delta} n^2 r^{m + 2} |f|_{H^1(\Xset)}^2
+	\end{equation*}
+\end{proposition}
+
+In Proposition~\ref{prop:calder19_1}, it is assumed that $r$, $\wt{\delta}$ and $\theta$ satisfy the following smallness conditions.
+\begin{enumerate}[label=(S\arabic*)]
+	\item 
+	\setcounter{enumi}{1}
+	\begin{equation*}
+	n^{-1/m} < \wt{\delta} \leq \frac{1}{4}r~~\textrm{and}~~C(\theta + \wt{\delta}) \leq \frac{1}{2}p_{\min}~~\textrm{and}~~r\leq\min\{c_3,1\}.
+	\end{equation*}
+\end{enumerate}
+
+\begin{proposition}[c.f Theorem 2.4 of~\citep{calder2019}]
+	\label{prop:calder19_1}
+	With probability at least $1 - Cn\exp(-cn\theta^2\wt{\delta}^m)$, the following statement holds. For any $k \in \mathbb{N}$ such that
+	\begin{equation*}
+	\sqrt{\lambda_k(\Delta_P)}r + C(\theta + \wt{\delta}) \leq \frac{1}{2}
+	\end{equation*} 
+	it holds that
+	\begin{equation*}
+	nr^{m+2} \lambda_k(\Delta_P) \biggl(1 - C\Bigl(r(\sqrt{\lambda_k(\Delta_P)} + 1) + \frac{\wt{\delta}}{r} + \theta\Bigr)\biggr)\leq \lambda_k(G_{n,r}) \leq nr^{m+2} \lambda_k(\Delta_P) \biggl(1 + C\Bigl(r(\sqrt{\lambda_k(\Delta_P)} + 1) + \frac{\wt{\delta}}{r} + \theta\Bigr)\biggr)
+	\end{equation*}
+\end{proposition}
+
+Proposition~\ref{prop:calder19_2} follows from Lemma 3.1 of \citep{calder2019}, along with a union bound.
+\begin{proposition}
+	\label{prop:calder19_2}
+	With probability at least $1 - 2Cn\exp(-cp_{\max}nr^m)$, it holds that
+	\begin{equation*}
+	D_{\max}(G_{n,r}) \leq Cnr^{m}.
+	\end{equation*}
+\end{proposition}
+
+Finally, we note that a Weyl's Law holds for Riemmanian manifolds without boundary, i.e.
+\begin{equation*}
+\lambda_k(\Delta_P) \asymp k^{2/m}.
+\end{equation*}
+Put $B_{n,r}(k) := \min\{nr^{m + 2}k^{2/m}, nr^m\}$.
+Following parallel steps to the proof of Lemma~\ref{lem:neighborhood_eigenvalue}, one can derive from Propositions~\ref{prop:calder19_1} and~\ref{prop:calder19_2}, and Weyl's Law, that with probability at least $1 - Cn\exp(-cnr^m)$,
+\begin{equation}
+\label{eqn:neighborhood_eigenvalue_manifold}
+cB_{n,r}(k) \leq \lambda_k \leq CB_{n,r}(k),~~\textrm{for all $2 \leq k \leq n$}
+\end{equation}
+
+\section{Proofs of Main Results}
 \label{sec:main_results}
-
-\subsection{Proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}}
+We are now in a position to prove Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing_manifold}, as well as a few other claims from our main text.
+We divide our proofs into estimation results and testing results.
+Throughout, it will be convenient for us to deal with the normalization $\wt{\rho} = \rho nr^{d + 2}$.
+
+\subsection{Proof of estimation results}
 \label{subsec:laplacian_smoothing_estimation1_pf}
-Let $\mc{E}$ be the set of $\mathbf{X} = (X_1,\ldots,X_n)$ for which
-\begin{enumerate}[(i)]
-	\item
-	\label{pf:laplacian_smoothing_estimation1_0}
-	$f_0^T \Lap_{n,r} f_0 \leq \frac{p_{\max}^2 \sigma_K}{\delta} n^2 r^{d + 2}M^2$
-	\item 
-	\label{pf:laplacian_smoothing_estimation1_1}
-	$c_3 \cdot \min\{k^{2/d}nr^{d+2},nr^d\} \leq \lambda_k(G_{n,r}) \leq C_3k^{2/d}nr^{d + 2} $ for all $2 \leq k \leq n$.
-\end{enumerate}
-We already know, from Lemma~\ref{lem:graph_sobolev_seminorm} and Corollary~\ref{cor:neighborhood_eigenvalue}, that for any $\delta > 0$ and $(\log(n)/n)^{1/d} \leq r \leq a_0$,
-\begin{equation}
-\label{pf:laplacian_smoothing_estimation1_1.5}
-\Pbb\bigl({\bf X} \in \mc{E}\bigr) \geq 1 - \delta - A_1n\exp\bigl(-a_1nr^d\bigr).
-\end{equation}
-We will now show that conditional on $\mathbf{X} \in \mc{E}$, the empirical mean square error is small with high probability.
-Having conditioned on $\mathbf{X}$, we can use Lemma~\ref{lem:ls_fixed_graph_estimation} to control the remaining randomness arising from the noisy response $Y$, which in this context implies
-\begin{equation}
-\label{pf:laplacian_smoothing_estimation1_2}
-\begin{aligned}
-\Pbb\Biggl(& \frac{1}{n}\bigl\|\wt{f}_{\LS}(G_{n,r}) - f_0\bigr\|_2^2 \leq \frac{2\rho}{n} \bigl(f_0^T \Lap_{n,r} f_0\bigr) + \frac{10}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k}(G_{n,r}) + 1\bigr)^2} \Bigg| \mathbf{X} = x \in \mc{E}\Biggr) \\
-& \geq 1 - \exp\Biggl\{-\sum_{k = 1}^{n}\frac{1}{\bigl(\rho \lambda_k(G_{n,r}) + 1\bigr)^2}\Biggr\} \\
-& \geq 1 - \exp\Biggl\{-\sum_{k = 1}^{n}\frac{1}{\bigl(C_3M^{-4/(2+d)}k^{2/d}n^{-2/(2+d)} + 1\bigr)^2}\Biggr\} \\
-& \geq 1 - \exp\Bigl\{-c M^{d/(2d + 4)} n^{d/(2+d)}\Bigr\}
-\end{aligned}
-\end{equation}
-where the second inequality follows from the upper bound in~\ref{pf:laplacian_smoothing_estimation1_1} and the choice of regularization parameter $\rho = M^{-4/(2+d)}(nr^{d+2})^{-1}n^{-2/(2+d)}$.
-
-It remains only to upper bound the (conditional on ${\bf X} = x$ for some $x \in \mc{E}$) squared bias and variance terms.
-A sufficient upper bound on the bias term comes immediately from the upper bound~\ref{pf:laplacian_smoothing_estimation1_0} and our choice of $\rho$,
-\begin{equation}
-\label{pf:laplacian_smoothing_estimation1_2.5}
-\frac{\rho}{n} \bigl(f_0^T \Lap_{{n,r}} f_0\bigr) \leq \frac{p_{\max}^2 \sigma_K}{\delta} M^{2/(2+d)} n^{-2/(2 + d)}.
-\end{equation}
-
-To upper bound the variance term, we replace the eigenvalues $\lambda_k(G_{n,r})$ by their lower bounds in~\ref{pf:laplacian_smoothing_estimation1_1} and plug in our choice of $\rho$ to obtain,
+
+\paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}.}
+
+From Lemmas~\ref{lem:graph_sobolev_seminorm},~\ref{lem:neighborhood_eigenvalue}, and~\ref{lem:ls_fixed_graph_estimation}, we have that with probability at least $1 - \delta - C_1n\exp(-c_1nr^d) - \exp(-\sum_{k = 1}^{n}(\rho \lambda_k + 1)^{-2})$,
 \begin{align}
-\frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k}(G_{n,r}) + 1\bigr)^2} & \leq \frac{C}{n}\sum_{k = 1}^{n} \biggl\{\frac{1}{(\rho k^{2/d} n r^{d + 2} + 1)^2} + \frac{1}{(\rho n r^{d} + 1)^2} \biggr\} \nonumber \\
-& \leq \frac{C}{n}\sum_{k = 1}^{n} \biggl\{\frac{1}{\bigl(M^{-4/(2 + d)}n^{-2/(2+d)}k^{2/d} + 1\bigr)^2}\biggr\} + Cn^{4/(2 + d)}M^{8/(2+d)}r^{4} \nonumber \\
-& \leq CM^{2d/(2 + d)} n^{-2/(2 + d)} + CM^{8/(2 + d)}n^{(2 - d)/(2 + d)} \sum_{k = k_{\star}}^{n} \Bigl\{\frac{1}{k^{4/d}}\Bigr\}  + CM^{8/(2 + d)}n^{4/(2 + d)}r^4 \label{pf:laplacian_smoothing_estimation1_2.25}
+\|\wh{f} - f_0\|_n^2 & \leq \frac{C_3}{\delta} \wt{\rho} M^2 + \frac{10}{n} \sum_{k = 1}^{n} \Biggl(\frac{1}{\wt{\rho}\min\{k^{2/d},r^{-2}\} + 1}\Biggr)^2 \nonumber \\
+& \leq \frac{C_3}{\delta} \wt{\rho} M^2 + \frac{10}{n} \sum_{k = 1}^{n} \biggl(\frac{1}{\wt{\rho}k^{2/d} + 1}\biggr)^2 + \frac{10r^4}{\wt{\rho}^2} \label{pf:laplacian_smoothing_estimation1_1}
 \end{align}
-where $k_{\star} = M^{2d/(2 + d)}n^{d/(2 + d)}$.
-
-We now upper bound the 2nd and 3rd term on the right hand side of~\eqref{pf:laplacian_smoothing_estimation1_2.25}.
-To control the 3rd term, we use the upper bound $r \leq n^{-3/(4 + 2d)} M^{(d - 4)/(4 + 2d)}$ assumed in~\ref{asmp:ls_kernel_radius_estimation} to determine that
-\begin{equation}
-\label{pf:laplacian_smoothing_estimation1_3}
-M^{8/(2 + d)}n^{4/(2 + d)}r^4 \leq M^{2d/(2 + d)}n^{-2/(2 + d)} 
-\end{equation}
-To control the 2nd term on the right hand side of~\eqref{pf:laplacian_smoothing_estimation1_2.25}, we treat it as a Riemann sum evaluated at the right end points of $[k_{\star},k_{\star} + 1], [n - 1,n]$.
-We use the fact that such a Riemann sum upper bounds the integral of a montonically nonincreasing function to conclude that
-\begin{equation}
-\label{pf:laplacian_smoothing_estimation1_4}
-\begin{aligned}
-\sum_{k = k_{\star}}^{n} \frac{1}{k^{4/d}} & \leq \int_{k_{\star}}^{n} \frac{1}{x^{4/d}} \,dx \\
-& = - \frac{1}{(4/d - 1)} x^{-(4/d - 1)} \Big|_{k_{\star}}^{n} \\
-& \leq \frac{1}{(4/d - 1)} k_{\star}^{-(4/d - 1)}
-\end{aligned}
-\end{equation}
-where the equality in the above computation holds because $d < 4$.
-Recalling the definition of $k_{\star}$, plugging this back into~\eqref{pf:laplacian_smoothing_estimation1_3} implies a sufficient upper bound on the variance term.
-Along with~\eqref{pf:laplacian_smoothing_estimation1_1.5}-\eqref{pf:laplacian_smoothing_estimation1_2.5}, this establishes the claim.
-
-\paragraph{Proving near-optimal bounds when $d = 4$.}
-Note that nothing in the derivations of  ~\eqref{pf:laplacian_smoothing_estimation1_1.5}-\eqref{pf:laplacian_smoothing_estimation1_2.5} relied on $d < 4$, and so they all hold when $d = 4$.
-We now upper bound the 2nd and 3rd terms in~\eqref{pf:laplacian_smoothing_estimation1_2.5}.
-By setting $r = C_4(\log(n)/n)^{1/4}$ for some constant $C_4$, we have that
-\begin{equation*}
-M^{8/(2+d)}n^{4/(2+d)}r^4 = C_4^4 M^{8/(2+d)} \log(n) n^{-1/3},
-\end{equation*}
-taking care of the third term. On the other hand, we have that
-\begin{equation*}
-\sum_{k = k_{\star}}^{n} \frac{1}{k} \leq \log(n)
-\end{equation*}
-and so the second term is also upper bounded by $C M^{8/(2+d)} \log(n) n^{-1/3}$.
-
-\subsection{Proof of Theorem~\ref{thm:laplacian_smoothing_testing}}
-In this proof we will need to invoke Lemma~\ref{lem:empirical_norm_sobolev}, and it is in order to satisfy the conditions of this Lemma that we require $M \leq n^{(4 - d)/(4 + d)}$.
-More specifically, by~\eqref{eqn:laplacian_smoothing_testing} along with the restriction $M \leq n^{(4 - d)/(4 + d)}$,
-\begin{equation*}
-\norm{f}_{\Leb^2(\Xset)} \geq \frac{C}{\delta^2} M^{2d/(4 + d)} n^{-4/(4 + d)} \geq \frac{C}{\delta^2} M^2 n^{-2/d}
-\end{equation*}
-and~\eqref{eqn:empirical_norm_sobolev_1} is therefore satisfied when we choose $C \geq C_6$ in~\eqref{eqn:laplacian_smoothing_testing}.
-
-We proceed to prove Theorem~\ref{thm:laplacian_smoothing_testing}.
-Throughout this proof we set $\delta = 1/b$.
-We will proceed, as in Subsection~\ref{subsec:laplacian_smoothing_estimation1_pf}, by conditioning on the random design $\mathbf{X}$.
-Let $\mc{E}(f_0)$ be the set of $\mathbf{X} \in \Xset^n$ such that~\eqref{eqn:graph_sobolev_seminorm},~\eqref{eqn:neighborhood_eigenvalue_2}, and~\eqref{eqn:empirical_norm_sobolev_1} hold: by Lemmas~\ref{lem:graph_sobolev_seminorm} and~\ref{lem:empirical_norm_sobolev}, and Corollary \ref{cor:neighborhood_eigenvalue},
-\begin{equation*}
-\Pbb\bigl(\mc{E}(f_0)\bigr) \geq 1 - 6\delta - A_1n\exp\bigl\{-a_1nr^d\bigr\}. 
-\end{equation*}
-For any $\mathbf{X} \in \mc{E}(f_0)$ the conditional variance of $T_{\LS}$ can be upper bounded using~\eqref{eqn:neighborhood_eigenvalue_2} and~\ref{asmp:ls_kernel_radius_testing},
-\begin{align*}
-\frac{1}{n} \biggl(\sum_{k = 1}^{n} \bigl[\lambda_k(G_{n,r}) + 1\bigr]^{-4}\biggr)^{1/2} & \leq \frac{1}{c_5^2 n} \biggl(M^{\frac{32}{8 + d}}r^8n^{\frac{20+d}{4+d}} + \sum_{k = 1}^{n} \bigl[n^{-\frac{4}{4 + d}}k^{\frac{2}{d}}M^{-\frac{8}{4 + d}} + 1\bigr]^{-4}\biggr)^{1/2} \\
-& \leq \frac{1}{c_5^2 n} \biggl(M^{\frac{4d}{4 + d}}n^{\frac{2d}{4 + d}} + \sum_{k = 1}^{n} \bigl[n^{-\frac{4}{4 + d}}k^{\frac{2}{d}}M^{-\frac{8}{4 + d}} + 1\bigr]^{-4}\biggr)^{1/2} \\
-& \leq \frac{1}{c_5^2 n} \biggl(3 M^{\frac{4d}{4 + d}}n^{2d/(4 + d)}\biggr)^{1/2} \\
-& = \frac{\sqrt{3}}{c_5^2} M^{\frac{2d}{4 + d}} n^{-\frac{4}{4 + d}}.
-\end{align*}
-The last inequality in the above display similarly to~\eqref{pf:laplacian_smoothing_estimation1_4}. Setting $k_{\star} = n^{2d/(4+d)}M^{4d/(4 + d)}$, we derive that
-\begin{equation}
-\begin{aligned}
-\label{pf:laplacian_smooting_testing1}
-\sum_{k = 1}^{n} \bigl[n^{-\frac{4}{4 + d}}k^{\frac{2}{d}}M^{-\frac{8}{4 + d}} + 1\bigr]^{-4} & \leq k_{\star} + n^{\frac{16}{4 + d}} M^{\frac{32}{4 + d}}\sum_{k = k_{\star}}^{n} k^{-\frac{8}{d}}\\
-& \leq k_{\star} + n^{\frac{16}{4 + d}} M^{\frac{32}{4 + d}} \int_{k_{\star}}^{n} x^{-\frac{8}{d}} \,dx \\
-& \leq  k_{\star} + \frac{1}{(8/d - 1)}n^{\frac{16}{4 + d}} M^{\frac{32}{4 + d}} k_{\star}^{1 - 8/d} \\
-& \leq 2k_{\star}.
-\end{aligned}
-\end{equation}
-As a result, 
-\begin{align*}
-\frac{2\rho}{n} \bigl(f_0^T \Lap_{n,r} f_0) + \frac{4}{\delta n}\biggl(\sum_{k = 1}^{n} \bigl[\lambda_k(G_{n,r}) + 1\bigr]^{-4}\biggr)^{1/2} & \leq \frac{1}{\delta}\biggl(2C_1 + \frac{4\sqrt{3}}{c_5^2}\biggr)n^{-\frac{4}{4 + d}} M^{\frac{2d}{4 + d}} \\
-& \leq \delta p_{\min} \norm{f}_{\Leb^2(\Xset)}^2 \leq \delta \Ebb\bigl[\norm{f}_n^2\bigr] \\
-& \leq \norm{f}_n^2,
-\end{align*}
-where the second inequality follows upon choosing $C \geq p_{\min}^{-1}(2C_1 + 4\sqrt{3}/c_5^2)$ in~\eqref{eqn:laplacian_smoothing_testing}.
-We may therefore bound the Type II error of $\wt{\phi}_{\LS}$ by appealing to Lemma~\ref{lem:ls_fixed_graph_testing},
-\begin{align*}
-\Ebb_{f_0}\Bigl[1 - \wt{\phi}_{\LS}\Bigr] & \leq \Ebb_{f_0}\Bigl[1 - \wt{\phi}_{\LS}\big| \mathbf{X} \in \mc{E}(f_0)\Bigr] + 1 - \Pbb\Bigl(\mc{E}(f_0)\Bigr) \\
-& \leq 4 \delta^2 + 8 \delta \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G_{n,r}) + 1)^4} \Biggr)^{-1/2} + 6\delta + A_1\exp\bigl\{-a_1nr^{d}\bigr\} \\
-& \leq 4 \delta^2 + 16\delta n^{-d(4 + d)}M^{-2d/(4 + d)}  + 6\delta + A_1\exp\bigl\{-a_1nr^{d}\bigr\}
-\end{align*}
-where the last inequality follows from the reversing the inequalities in~\eqref{pf:laplacian_smooting_testing1}, with a factor of $1/2$ instead of $2$ in front.
-
-\subsection{Proof of~\eqref{eqn:laplacian_smoothing_testing_low_smoothness}}
-When $\rho = 0$, the Laplacian smoother $\wh{f}_{\LS} = Y$, the test statistic $T_{\LS} = \frac{1}{n}\|Y\|_2^2$, and the threshold $\wh{t}_b = 1 + 2bn^{-1/2}$. The expectation of $T_{\LS}$ is 
-\begin{equation*}
-\Ebb\bigl[T_{\LS}\bigr] = \mathbb{E}\bigl[f_0^2(X)\bigr] + 1 \geq p_{\min} \norm{f_0}_{\Leb^2(\Xset)}^2 + 1
-\end{equation*}
-When $f_0 \in \Leb^4(\Xset,M)$, the variance can be upper bounded
-\begin{equation*}
-\Var\bigl[T_{\LS}\bigr] \leq \frac{1}{n}\Bigl(3 + p_{\max} M^4 + p_{\max}\norm{f_0}_{\Leb^2(\Xset)}^2\Bigr).
-\end{equation*}
-Now, let us assume that
-\begin{equation*}
-\norm{f_0}_{\Leb^2(X)}^2 \geq \frac{4b}{p_{\min}} n^{-1/2},
-\end{equation*}
-so that $E[T_{\LS}] - \wh{t}_b \geq E[T_{\LS}]/2$. Hence, by Chebyshev's inequality
-\begin{align*}
-\mathbb{P}_{f_0}\Bigl(T_{\LS} \leq \wh{t}_b \Bigr) & \leq 4 \frac{\Var_{f_0}\bigl[T_{\LS}\bigr]}{\mathbb{E}[f^2(X)]^2} \\
-& \leq \frac{4}{n} \cdot \frac{ 3 + p_{\max}\bigl(M^4 + \|f_0\|_{\Leb^2(\Xset)}^2 \bigr)}{p_{\min}^2 \|f_0\|_{\Leb^2(\Xset)}^4} \\
-& \leq \frac{1}{16b^2}\Bigl(3 + p_{\max}n^{-1/2} + p_{\max}M^4\Bigr).
-\end{align*}
-\subsection{Bounds on $\Leb^2(\Xset)$ error under Lipschitz assumption.}
+where $C_3 := (1 + L_pc_0)^2 \sigma_K$.
+Setting $\wt{\rho} = M^{-4/(2 + d)}n^{-2/(2 + d)}$ as prescribed in Theorem~\ref{thm:laplacian_smoothing_estimation1}, from Lemma~\ref{lem:variance_term_estimation}:
+\begin{align*}
+\|\wh{f} - f_0\|_n^2 & \leq\frac{C_3}{\delta} \wt{\rho} M^2 + \frac{20\wt{\rho}^{-d/2}}{n}  + \frac{10r^4}{\wt{\rho}^2} \\
+& = \biggl(\frac{C_3}{\delta} + 10\biggr)M^{2d/(2+d)} n^{-2/(2 + d)} + 10r^4M^{8/(2 + d)}n^{4/(2 + d)} \\
+& \leq \biggl(\frac{C_3}{\delta} + 20\biggr)M^{2d/(2+d)} n^{-2/(2 + d)}
+\end{align*}
+with the last inequality following from~\ref{asmp:ls_kernel_radius_estimation}. Finally, observe that by Lemmas~\ref{lem:neighborhood_eigenvalue} and~\ref{lem:variance_term_estimation},
+\begin{equation*}
+\exp\biggl(-\sum_{k = 1}^{n}(\rho \lambda_k + 1)^{-2}\biggr) \leq \exp\biggl(-\sum_{k = 1}^{n}(\wt{\rho} k^{2/d} + 1)^{-2}\biggr) \leq \exp\biggl(-\frac{1}{4}\min\Bigl\{\floor{\wt{\rho}^{-d/2}},n\Bigr\}\biggr),
+\end{equation*}
+completing the proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}.
+
+\paragraph{Nearly-optimal rates when $d = 4$.}
+Continuing on from~\eqref{pf:laplacian_smoothing_estimation1_1}, from Lemma~\ref{lem:variance_term_estimation} we have that
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 \leq \frac{C_3}{\delta} \wt{\rho} M^2 + \frac{10}{n\wt{\rho}^2} +  \frac{10 \log n}{n\wt{\rho}^2} + \frac{10r^4}{\wt{\rho}^2}
+\end{equation*}
+Setting $r = (C_4\log(n)/n)^{1/4}$, we obtain
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 \leq \frac{C_3}{\delta} \wt{\rho} M^2 + \frac{10}{n\wt{\rho}^2} +  \frac{10 \log n}{n\wt{\rho}^2} + \frac{10C_4\log n}{n\wt{\rho}^2},
+\end{equation*}
+and choosing $\wt{\rho} = M^{-2/3}(\log n/n)^{1/3}$ yields
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 \leq \biggl(\frac{C_3}{\delta} + 20 + 10C_4\biggr) M^{4/3} \biggl(\frac{\log n}{n}\biggr)^{1/3}.
+\end{equation*}
+
+\paragraph{Suboptimal rates when $d > 4$.}
+Once again continuing on from~\eqref{pf:laplacian_smoothing_estimation1_1}, from Lemma~\ref{lem:variance_term_estimation} we have that,
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 \leq \frac{C_3}{\delta} \wt{\rho} M^2 + \frac{10}{n\wt{\rho}^{d/2}} +  \frac{10}{n^{4/d}\wt{\rho}^2} + \frac{10r^4}{\wt{\rho}^2}
+\end{equation*}
+Setting $r = (C_4\log n/n)^{1/d}$, we obtain
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 \leq \frac{C_3}{\delta} \wt{\rho} M^2 + \frac{10}{n\wt{\rho}^{d/2}} +  \frac{10}{n^{4/d}\wt{\rho}^2} + \frac{10C_4(\log n)^{4/d}}{n^{4/d}\wt{\rho}^2},
+\end{equation*}
+and choosing $\wt{\rho} = M^{-2/3}n^{-4/(3d)}$ yields
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 \biggl(\frac{C_3}{\delta} + 10 + 10C_4\biggr)M^{4/d}\biggl(\frac{\log(n)}{n^{1/3}}\biggr)^{4/d} + 10M^{d/3}n^{-1/3}.
+\end{equation*}
+
+\paragraph{Bounds on $\Leb^2(\Xset)$ error under Lipschitz assumption.}
 Suppose $f_0$ has bounded Lipschitz constant.
@@
 \begin{equation*}
-\wh{f}(x) := \sum_{i = 1}^{n} \cdot \wh{f}_i \1\{x \in V_i\}.
+\wh{f}(x) := \sum_{i = 1}^{n} \wh{f}_i \cdot \1\{x \in V_i\}.
 \end{equation*}
@@
 	\label{prop:out_of_sample_error}
-	Suppose $f_0$ satisfies $|f_0(x') - f_0(x)| \leq M \|x' - x\|$ for all $x',x \in \mc{X}$. Then, with probability at least $1 - \delta$ it holds that
+	Suppose $f_0$ satisfies $|f_0(x') - f_0(x)| \leq M \|x' - x\|$ for all $x',x \in \mc{X}$. Then for all $n$ sufficiently large, with probability at least $1 - \delta$ it holds that
 	\begin{equation*}
@@
 Note that $n^{-2/d} \ll n^{-2/(2 +d)}$.
-Therefore Proposition~\ref{prop:out_of_sample_error} together with Theorem~\ref{thm:laplacian_smoothing_estimation1} imply that with high probability, $\wh{f}$ achieves the nearly-optimal (up to a factor of $\log n$) estimation out-of-sample error---that is, $\|\wh{f} - f_0\|_{\Leb^2(\Xset)}^2 \leq C \log(n) M^{2d/(2 + d)}n^{-2/(2+d)}$---as long as $M \leq Cn^{1/d}$.
+Therefore Proposition~\ref{prop:out_of_sample_error} together with Theorem~\ref{thm:laplacian_smoothing_estimation1} implies that with high probability, $\wh{f}$ achieves the nearly-optimal (up to a factor of $\log n$) estimation out-of-sample error---that is, $\|\wh{f} - f_0\|_{\Leb^2(\Xset)}^2 \leq C \log(n) M^{2d/(2 + d)}n^{-2/(2+d)}$---as long as $M \leq Cn^{1/d}$.
 \paragraph{Proof of Proposition~\ref{prop:out_of_sample_error}.}
@@
 
-Now we will use some results of \citep{chaudhuri2010} regarding uniform concentration of empirical counts. Letting
-\begin{equation*}
-\varepsilon_n := \biggl(\frac{2C_{o}\log(1/\delta)d\log n}{\nu_dp_{\min}a_0n}\biggr)^{1/d}
-\end{equation*}
-we have that for every $x \in \Xset$, $P(B(x,\varepsilon_n)) \geq 2C_{o}\log(1/\delta)d\frac{\log n}{n}$.
+Now we will use some results of \citep{chaudhuri2010} regarding uniform concentration of empirical counts, to upper bound $\diam(V_i)$ Set
+\begin{equation*}
+\varepsilon_n := \biggl(\frac{2C_{o}\log(1/\delta)d\log n}{\nu_dp_{\min}a_3n}\biggr)^{1/d}
+\end{equation*}
+and note that for $n$ sufficiently large, $\varepsilon_n \leq c_0$.
+Thus by~\eqref{eqn:integral_boundary} we have that for every $x \in \Xset$, $P(B(x,\varepsilon_n)) \geq 2C_{o}\log(1/\delta)d\frac{\log n}{n}$.
 Consequently, by Lemma~16 of \citep{chaudhuri2010} it holds that with probability at least $1 - \delta$,
@@
 \label{pf:out_of_sample_error_1}
-\textrm{for all $x \in \mc{X}$},~~ B(x,\varepsilon) \cap \{X_1,\ldots,X_n\} \neq \emptyset.
+\textrm{for all $x \in \mc{X}$},~~ B(x,\varepsilon_n) \cap \{X_1,\ldots,X_n\} \neq \emptyset.
 \end{equation}
@@
 
-\textcolor{red}{(TODO)}: Why do we have the lower bound on $P(B(x,\varepsilon_n))$?.
+\paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_estimation_manifold}.}
+The proof of Theorem~\ref{thm:laplacian_smoothing_estimation_manifold} follows exactly the same steps as the proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}, replacing the references to Lemma~\ref{lem:graph_sobolev_seminorm} and~\ref{lem:neighborhood_eigenvalue} by references to Propositions~\ref{prop:garciatrillos19_1} and~\eqref{eqn:neighborhood_eigenvalue_manifold}, and the ambient dimension $d$ by the intrinsic dimension $m$.
+
+\subsection{Proofs of testing results}
+
+\paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_testing}.}
+We would like to invoke Lemma~\ref{lem:ls_fixed_graph_testing}, and in order to do so, we must show that~\eqref{eqn:ls_fixed_graph_testing_critical_radius} is satisfied with respect to $G = G_{n,r}$.
+On the one hand, by Lemmas~\ref{lem:graph_sobolev_seminorm} and~\ref{lem:neighborhood_eigenvalue}, with probability at least $1 - b - C_1n\exp(-c_1nr^d)$,
+\begin{align*}
+\frac{2 \rho}{n} \bigl(f_0^T \Lap f_0\bigr) + \frac{4b}{n} \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{1/2} & \leq C_3b\wt{\rho}M^2 + \frac{4b}{n}\Biggl[\biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{1/2} + \frac{r^8}{\wt{\rho}^4n^{1/2}}\Biggr] \\
+& \leq C_3b\wt{\rho}M^2 + \frac{4b}{n}\biggl(\sqrt{2}\wt{\rho}^{-d/4} + \frac{r^8}{\wt{\rho}^4n^{1/2}}\biggr) \\ 
+& \leq \Bigl(C_3 + 12\Bigr)bM^{2d/(4 + d)} n^{-4/(4 + d)}
+\end{align*}
+where the second inequality comes from Lemma~\ref{lem:variance_term_testing}, and in the third inequality we set $\wt{\rho} = M^{-8/(4 + d)}n^{-4/(4 + d)}$ as prescribed by Theorem~\ref{thm:laplacian_smoothing_testing}.
+
+On the other hand, by Lemma~\ref{lem:empirical_norm_sobolev}, with probability at least $1 - 5/b$,
+\begin{equation*}
+\|f\|_n^2 \geq \frac{p_{\min}}{b} \|f\|_{\Leb^2(\Xset)}^2 \geq CbM^{2d/(4 + d)}n^{-4/(4 + d)}
+\end{equation*}
+and so choosing $C$ appropriately in~\eqref{eqn:laplacian_smoothing_testing}, ~\eqref{eqn:ls_fixed_graph_testing_critical_radius} is verified.
+
+By these derivations along with Lemma~\ref{lem:ls_fixed_graph_testing}, the Laplacian smoothing test $\wh{\varphi}$ has Type II error upper bounded by
+\begin{align*}
+\Pbb_{f_0}\bigl(\wh{T} \leq \wh{t}_b\bigr) & \leq \frac{6}{b} + \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2} + C_1n\exp(-c_1nr^d)\\
+& \leq \frac{6}{b} + \frac{4}{b^2} + \frac{32}{b} \max\Bigl\{\floor{\wt{\rho}^{-d/4}},n^{-1/2}\Bigr\} + C_1n\exp(-c_1nr^d)
+\end{align*}
+establishing the claim of Theorem~\ref{thm:laplacian_smoothing_testing}.
+
+\paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_testing_manifold}.}
+The proof of Theorem~\ref{thm:laplacian_smoothing_testing_manifold} follows exactly the same steps as the proof of Theorem~\ref{thm:laplacian_smoothing_testing}, replacing the references to Lemma~\ref{lem:graph_sobolev_seminorm} and~\ref{lem:neighborhood_eigenvalue} by references to Propositions~\ref{prop:garciatrillos19_1} and~\eqref{eqn:neighborhood_eigenvalue_manifold}, and the ambient dimension $d$ by the intrinsic dimension $m$.
+
+\paragraph{Proof of~\eqref{eqn:laplacian_smoothing_testing_low_smoothness}.}
+When $\rho = 0$, the Laplacian smoother $\wh{f} = Y$, the test statistic $\wh{T} = \frac{1}{n}\|Y\|_2^2$, and the threshold $\wh{t}_b = 1 + 2bn^{-1/2}$. The expectation of $\wh{T}$ is 
+\begin{equation*}
+\Ebb\bigl[\wh{T}\bigr] = \mathbb{E}\bigl[f_0^2(X)\bigr] + 1 \geq p_{\min} \norm{f_0}_{\Leb^2(\Xset)}^2 + 1
+\end{equation*}
+When $f_0 \in \Leb^4(\Xset,M)$, the variance can be upper bounded
+\begin{equation*}
+\Var\bigl[\wh{T}\bigr] \leq \frac{1}{n}\Bigl(3 + p_{\max} M^4 + p_{\max}\norm{f_0}_{\Leb^2(\Xset)}^2\Bigr).
+\end{equation*}
+Now, let us assume that
+\begin{equation*}
+\norm{f_0}_{\Leb^2(X)}^2 \geq \frac{4b}{p_{\min}} n^{-1/2},
+\end{equation*}
+so that $E[\wh{T}] - \wh{t}_b \geq E[\wh{T}]/2$. Hence, by Chebyshev's inequality
+\begin{align*}
+\mathbb{P}_{f_0}\Bigl(\wh{T} \leq \wh{t}_b \Bigr) & \leq 4 \frac{\Var_{f_0}\bigl[\wh{T}\bigr]}{\mathbb{E}[f^2(X)]^2} \\
+& \leq \frac{4}{n} \cdot \frac{ 3 + p_{\max}\bigl(M^4 + \|f_0\|_{\Leb^2(\Xset)}^2 \bigr)}{p_{\min}^2 \|f_0\|_{\Leb^2(\Xset)}^4} \\
+& \leq \frac{1}{16b^2}\Bigl(3 + p_{\max}n^{-1/2} + p_{\max}M^4\Bigr).
+\end{align*}
+
+\subsection{Two convenient estimates}
+The following Lemma provides convenient upper and lower bounds on our estimation variance term.
+\begin{lemma}
+	\label{lem:variance_term_estimation}
+	For any $t > 0$,
+	\begin{equation*}
+	\min\Bigl\{\frac{1}{4}\floor{t^{-d/2}},n\Bigr\} \leq \sum_{k = 1}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^2 \leq t^{-d/2} +
+	\begin{cases*}
+	t^{-d/2},& ~~\textrm{if $d < 4$} \\
+	\frac{1}{t^2}\log n,& ~~\textrm{if $d = 4$} \\
+	\frac{1}{t^2}n^{-(4/d - 1)},&~~\textrm{if $d > 4$.}
+	\end{cases*}
+	\end{equation*}
+\end{lemma}
+
+\begin{lemma}
+	\label{lem:variance_term_testing}
+	For any $t > 0$, when $d < 4$,
+	\begin{equation*}
+	\min\Bigl\{\frac{1}{16}\floor{t^{-d/2}},n\Bigr\} \leq \sum_{k = 1}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^4 \leq 2t^{-d/2}.
+	\end{equation*}
+\end{lemma}
+\paragraph{Proof of Lemma~\ref{lem:variance_term_estimation}.}
+We begin by proving the upper bounds. Observe that
+\begin{align*}
+ \sum_{k = 1}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^2 & \leq t^{-d/2} + \frac{1}{t^2}\sum_{k = t^{-d/2} + 1}^{n} \frac{1}{k^{4/d}} \\
+ & \leq t^{-d/2} + \frac{1}{t^2} \int_{t^{-d/2}}^{n} x^{-4/d} \,dx
+\end{align*}
+and the various upper bounds (for $d < 4$, $d = 4$, and $d > 4$) follow by doing a straightforward integral.
+
+For the lower bound, we simply recognize that when $k \leq \floor{t^{-d/2}}$,
+\begin{equation*}
+\biggl(\frac{1}{tk^{2/d} + 1}\biggr)^2 \geq \frac{1}{4}.
+\end{equation*}
+
+\paragraph{Proof of Lemma~\ref{lem:variance_term_testing}.}
+The upper bound follows similarly to that of Lemma~\ref{lem:variance_term_estimation}:
+\begin{align*}
+\sum_{k = 1}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^4 & \leq t^{-d/2} + \frac{1}{t^4}\sum_{k = t^{-d/2} + 1}^{n} \frac{1}{k^{8/d}} \\
+& \leq t^{-d/2} + \frac{1}{t^4} \int_{t^{-d/2}}^{n} x^{-8/d} \,dx \\
+& \leq 2t^{-d/2}.
+\end{align*}
+
+The lower bound follows from the same logic as we used to derive the lower bound in Lemma~\ref{lem:variance_term_estimation}.
 
@@
 	\end{equation*}
-	with probability at least $1 - 2n\exp\Bigl(-nr^da_0p_{\min}/(3[K(0)]^2)\Bigr)$. 
+	with probability at least $1 - 2n\exp\Bigl(-nr^da_3p_{\min}/(3[K(0)]^2)\Bigr)$. 
 \end{lemma}
@@
 & \geq n p_{\min} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx \\
-& =  nr^d p_{\min}\int_{\mc{B}_x} K\bigl(\|z\|\bigr) \,dz \\
-& \geq nr^d a_0 p_{\min}
-\end{align*}
-with the final inequality following from~\eqref{eqn:lb_expected_degree}, and the normalization $\int_{\Rd} K(\|z\|) \,dz = 1$. Similar derivations yield the upper bound
+& \geq n p_{\min} a_3 \int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx \\
+& \geq nr^d p_{\min}.
+\end{align*}
+with the second inequality following from~\eqref{eqn:integral_boundary}, and the final inequality from the normalization $\int_{\Rd} K(\|z\|) \,dz = 1$. Similar derivations yield the upper bound
 \begin{equation*}

### 2020-10-26 notes/weekly_notes/10.28.20.NotesForWeek.tex
Added a note outlining where we get stuck in our proof of a lower bound for Laplacian smoothing.

\documentclass{article}
\usepackage{amsmath, amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{verbatim}
\usepackage{fullpage}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}

% Make a widecheck symbol (thanks, Stack Exchange!)
\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{
	<5> <6> <7> <8> <9> <10>
	<10.95> <12> <14.4> <17.28> <20.74> <24.88>
	mathx10
}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareFontSubstitution{U}{mathx}{m}{n}
\DeclareMathAccent{\widecheck}{0}{mathx}{"71}
% widecheck made

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\iid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}
\newcommand{\lambdavec}{\boldsymbol{\lambda}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}
\newcommand{\Lap}{L}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}
\newcommand{\Id}{I}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Vset}{\mathcal{V}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\SM}{\mathrm{SM}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

%%% Theorem environments
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Note on: Lower bound for Laplacian smoothing on Geometric Graphs}
\author{Alden Green}
\date{\today}
\maketitle

Suppose we observe design points $X_1,\ldots,X_n$, with associated responses
\begin{equation*}
Y_i = f_0(X_i) + \varepsilon_i;
\end{equation*}
here each $X_i$ belongs to a domain $\Xset$, the regression function $f_0: \Xset \to \Reals$, and $\varepsilon_i \sim N(0,1)$ are independent Gaussian noise. Let $G = (\{1,\ldots,n\},W)$ be a geometric graph defined over the vertices $\{1,\ldots,n\}$, which we associate with the design points $X_1,\ldots,X_n$, where $W_{ij} = 0$ unless $\|X_i - X_j\| \leq r$.  Let $\Lap$ be the (combinatorial) Laplacian matrix associated with $G$. The Laplacian smoothing estimator $\wh{f}$ is defined as
\begin{equation*}
\wh{f} := \argmin_{f \in \Reals^n} \Bigl\{\sum_{i = 1}^{n} (Y_i - f_0(X_i))^2 + \rho f^T \Lap f\Bigr\}.
\end{equation*}
The accuracy of $\wh{f}$ is measured by the in-sample mean squared error
\begin{equation*}
\|\wh{f} - f_0\|_n^2 := \frac{1}{n} \sum_{i = 1}^{n} \bigl(\wh{f}_i - f_0(X_i)\bigr)^2.
\end{equation*}

Suppose we make the following~\emph{random design} assumptions.
\begin{enumerate}[label=(P\arabic*)]
	\item
	\label{asmp:domain}
	 The domain $\Xset$ is an open, connected set, with Lipschitz boundary.
	\item 
	\label{asmp:density}
	The design points $X_1,\ldots,X_n$ are independently drawn by a distribution $P$ defined on $\Xset$. The distribution $P$ admits a density $p$ with respect to Lebesgue measure, which is bounded away from $0$ and $\infty$,
	\begin{equation*}
	0 < \frac{1}{\alpha} \leq p(x) \leq \alpha < \infty,~~\textrm{for all $x \in \Xset$.}
	\end{equation*}
\end{enumerate}
\begin{enumerate}[label=(G\arabic*)]
	\item 
	\label{asmp:geometric_graph}
	The graph $G$ is an unweighted \emph{random geometric graph}: $W_{ij} = 1$ if $\|X_i - X_j\| \leq r$ and $0$ otherwise.
\end{enumerate}
\begin{enumerate}[label=(F\arabic*)]
	\item 
	\label{asmp:sobolev_norm}
	The regression function $f_0$ belongs to the unit-norm first-order Sobolev ball $H^1(\Xset,1)$, meaning
	\begin{equation}
	\label{eqn:sobolev_norm}
	\int_{\Xset} \|\nabla f_0(x)\|^2 \,dx \leq 1
	\end{equation}
\end{enumerate}
We have derived the following upper bounds on the in-sample mean squared error of $\wh{f}$: with probability at least $1 - \delta$,
\begin{equation*}
\|\wh{f} - f_0\|_n^2 \leq C_{\delta}
\begin{cases*}
n^{-2/(2 + d)},& ~~\textrm{if $d = 1,2,3$} \\
\Bigl(\log n/n\Bigr)^{1/3},&~~\textrm{if $d = 4$} \\
n^{-4/(3d)},& ~~\textrm{if $d \geq 5$.}
\end{cases*}
\end{equation*}
These upper bounds fail to match the known minimax rate $n^{-2/(2 + d)}$ when $d \geq 4$. Our question is: is it our estimator, or our theory, that is deficient when $d \geq 4$? Simulations, and theory from the related fixed design problem, suggest it is the estimator. In this note, I detail this fixed design theory, and show the difficulty in applying similar reasoning to the random design case.

\section{Lower bounds in the fixed design setup}
In the fixed design setting, we assume $X_1,\ldots,X_n$ are a priori fixed, so that the only randomness comes from the noise variables $\varepsilon_i$ (one can do a better job of formalizing this assumption, but for the moment I won't bother.) In the fixed design setup, we make different assumptions.
\begin{enumerate}[label=(P\arabic*)]
	\setcounter{enumi}{2}	
	\item  
	\label{asmp:grid}
	The design points form a grid over the $d$-dimensional unit cube $[0,1]^d$. Adopting multiindex notation for conciseness, we have that for $N = n^{1/d}$, the design points are the set
	\begin{equation*}
	\Bigl\{X_j := \frac{1}{N}(j_1,\ldots,j_d): j \in [N]^d\Bigr\}
	\end{equation*}
\end{enumerate}

\begin{enumerate}[label=(G\arabic*)]
	\setcounter{enumi}{1}
	\item 
	\label{asmp:lattice}
	The graph $G$ is an unweighted lattice graph, with $W_{ij} = 1$ if $\|i - j\|_1 = 1$, and $0$ otherwise.
\end{enumerate}

\begin{enumerate}[label=(F\arabic*)]
	\setcounter{enumi}{1}
	\item 
	\label{asmp:graph_sobolev_norm}
	The regression function $f_0: \{X_1,\ldots,X_n\} \to \Reals$ satisfies
	\begin{equation*}
	\frac{1}{n^2\wb{r}^{d + 2}} f_0^{\top} \Lap f_0 \leq 1
	\end{equation*}
	for $\wb{r} := n^{-1/d}$.
\end{enumerate}
(NB: the grid and lattice assumption are not essential. We could likely replace the lattice by a general geometric graph of radius $r$, replace $\wb{r}$ by $r$, and replace the grid assumption by a Weyl's-Law-type assumption on the resulting eigenvalues of $G$. But the above is simpler, and gets the main idea across.)

We shall prove the following when $d > 4$: For any $\rho > 0$, there exists a function $f_0 \in \Reals^n$ satisfying~\ref{asmp:graph_sobolev_norm} such that
\begin{equation}
\label{eqn:lb_mse}
\mathbb{E}\bigl[\|\wh{f} - f_0\|_n^2\bigr] \gtrsim n^{-4/(3d)} \gg n^{-2/(d + 2)}.
\end{equation}

\paragraph{Proof of~\eqref{eqn:lb_mse}.}
Our strategy to prove this claim parallels the development of the corresponding upper bound. In the fixed design setup, the expected in-sample mean squared error of $\wh{f}$ can be decomposed into squared bias and variance:
\begin{align*}
\Ebb\bigl[\|\wh{f} - f_0\|_n^2\bigr] & = \underbrace{\|\Ebb[\wh{f}] - f_0]\|_n^2}_{\mathbb{B}(\wh{f},f_0)} + \underbrace{\Ebb\bigl[\|\wh{f} - \Ebb[\wh{f}]\|_n^2\bigr]}_{\mathbb{V}(\wh{f})}
\end{align*}
and the task reduces to lower bounding of lower bounding mean squared error reduces to lower bounding the bias and variance terms.

In the following two subsections, we give lower bounds on $\mathbb{B}(\wh{f},f_0)$ and $\mathbb{V}(\wh{f})$. The bounds take different forms, depending on which of the following situations we are in: (i) there exists $k \in \{2,\ldots,n/2 - 1\}$ such that $\rho \lambda_k = 1$, (ii) $\rho \lambda_2 > 1$, or (iii) $\rho \lambda_n < 1$. If we are in the first situation, then by Lemmas~\ref{lem:lb_bias} and~\ref{lem:lb_variance} there exists a choice of $f_0$ (which we make explicit in the statement of Lemma~\ref{lem:lb_bias}) such that
\begin{equation*}
\Ebb\bigl[\|\wh{f} - f_0\|_n^2\bigr] \geq c\biggl(\wb{\rho} + \frac{1}{\wb{\rho}^2}n^{-4/d}\biggr)
\end{equation*}
and a simple computation tells us that the right hand side is always greater than $cn^{-4/(3d)}$. 
If we are in the second situation, then by Lemma~\ref{lem:lb_bias} $\Ebb\bigl[\|\wh{f} - f_0\|_n^2\bigr] \gtrsim 1$; if we are in the third situation, then by Lemma~\ref{lem:lb_variance} the same thing is true, and so we have proved the claim.

(Of course, we have not handled the case where $\rho\lambda_k < 1 < \rho\lambda_{k + 1}$, but in spirit this is basically situation (i)---it should only introduce some negligible rounding error which I am not going to bother calculating right now.)

\subsection{Lower bound on Bias}
Let $S := (I + \rho \Lap)^{-1}$, and note that Laplacian smoothing can be written as $\wh{f} = S Y_{1:n}$ (writing $Y_{1:n} = (Y_1,\ldots,Y_n)$), and the bias of Laplacian smoothing as
\begin{equation}
\label{eqn:bias}
\mathbb{B}(\wh{f},f_0) = \frac{1}{n}f_0^{\top}(I - S)^2 f_0
\end{equation}
For $\rho > 0$, set $\wb{\rho} := \rho \cdot n\wb{r}^{d + 2}$.

\begin{lemma}
	\label{lem:lb_bias}
	Fix tuning parameter $\rho > 0$. 
	\begin{itemize}
		\item Suppose there exists a number $k \in \{1,\ldots,n\}$ such that $\rho \lambda_k = 1$. Then, setting $f_0$ equal to $\sqrt{\frac{n^2 \wb{r}^{d + 2}}{\lambda_k}}v_k$, we have that $(n^2\wb{r}^{d + 2})^{-1}f_0^{\top}\Lap f_0 = 1$, and 
		\begin{equation}
		\label{eqn:lb_bias_1}
		\mathbb{B}(\wh{f},f_0) = \frac{1}{4}\wb{\rho}
		\end{equation}
		\item Suppose $\lambda_1 \rho > 1$. Then, setting $f_0$ equal to $\sqrt{\frac{n^2 \wb{r}^{d + 2}}{\lambda_k}}v_1$, we have that $(n^2\wb{r}^{d + 2})^{-1}f_0^{\top}\Lap f_0 = 1$, and 
		\begin{equation}
		\label{eqn:lb_bias_2}
		\mathbb{B}(\wh{f},f_0) \geq c.
		\end{equation}
	\end{itemize}
\end{lemma}
Note: to properly trade off bias and variance when $d < 4$, we set $\wb{\rho} = n^{-2/(2 + d)}$. In this case, $\rho \lambda_k = 1$ implies that $k \asymp n^{d/(d + 2)}$, so that our choice of $f_0$ which saturates the upper bound $f_0^{\top} \Lap f_0 \lesssim \wb{\rho}$ (i.e the choice of $f_0$ for which~\eqref{eqn:lb_bias_1} holds) is a harmonic function of frequency $n^{1/(d + 2)}$. In plain English: its very wiggly.

\subsection{Lower bound on Variance}
The variance term $\mathbb{V}(\wh{f})$ is simply
\begin{equation*}
\mathbb{V}(\wh{f}) = \Ebb\bigl[\varepsilon^{\top} S^2 \varepsilon\bigr] = \frac{1}{n}\sum_{j = 1}^{n} \frac{1}{(\rho \lambda_j + 1)^2}
\end{equation*}

\begin{lemma}
	\label{lem:lb_variance}
	Fix tuning parameter $\rho > 0$. Assume $d > 4$.
	\begin{itemize}
		\item Suppose there exists a number $k \in \{1,\ldots,n/2 - 1\}$ such that $\rho \lambda_k = 1$. Then
		\begin{equation}
		\label{eqn:lb_variance_1}
		\mathbb{V}(\wh{f}) \geq c\frac{1}{\wb{\rho}^2}n^{-4/d}
		\end{equation}
		for $\wb{\rho} = (nr^{d + 2})\rho$. 
		\item Suppose instead that $\rho \lambda_{n/2 - 1} < 1$. Then
		\begin{equation}
		\label{eqn:lb_variance_2}
		\mathbb{V}(\wh{f}) \geq \frac{1}{4}.
		\end{equation}
	\end{itemize}
\end{lemma}

\section{What's tricky about a lower bound in the random design setup?}
The simple answer is: the bias. That is, it should not be hard to prove a rough analog to Lemma~\ref{lem:lb_variance} under the random design assumptions~\ref{asmp:density},~\ref{asmp:domain} and~\ref{asmp:geometric_graph}. However, to prove Lemma~\ref{lem:lb_bias} we set $f_0$ as a function of the graph $G$ (more specifically, as proportional to an eigenvector $v_k$ of $L$). In the random design setup, this is an impermissible operation, as we must choose the ``bad'' function $f_0$ before the design points $X_1,\ldots,X_n$ are drawn. Now, we could choose $f_0$ proportional to an eigenfunction $f_k$ of an appropriately weighted continuum Laplace-Beltrami operator. However, we would then need to argue that $f_k \approx v_k$, for a value of $k$ which is growing (quite rapidly) with $n$.

\section{Proofs}

\paragraph{Proof of Lemma~\ref{lem:lb_bias}.}
Take the representation~\eqref{eqn:bias}, and note that $(I - S)^2$ admits the eigendecomposition
\begin{equation*}
(I - S)^2 = \sum_{j = 1}^{n} \biggl(1 - \frac{1}{1 + \rho\lambda_j}\biggr)^2 v_j v_j^{\top} = \rho^2 \sum_{j = 1}^{n}\frac{\lambda_j^2}{(1 + \rho \lambda_j)^2} v_j v_j^{\top}
\end{equation*}
As a result, if $\lambda_k\rho = 1$, then 
\begin{equation*}
v_k^{\top} (I - S)^2 v_k = \rho^2 \frac{\lambda_k^2}{(1 + \rho \lambda_k)^2} = \frac{1}{4}
\end{equation*}
and setting $f_0 = \sqrt{\frac{n\wb{r}^{d+2}}{\lambda_k}}v_k$, we obtain
\begin{equation*}
\frac{1}{n}f_0^{\top} (I - S)^2 f_0 = \frac{n\wb{r}^{d + 2}}{\lambda_k} v_k^{\top} (I - S)^2 v_k =  \frac{1}{4}n\wb{r}^{d + 2}\rho.
\end{equation*}
On the other hand, if $\lambda_1\rho > 1$, then
\begin{equation*}
v_1^{\top} (I - S)^2 v_1  = \rho^2 \frac{\lambda_1^2}{(1 + \rho \lambda_1)^2} \geq \frac{1}{4}
\end{equation*}
and setting $f_0 = \sqrt{\frac{n\wb{r}^{d+2}}{\lambda_1}}v_1$ we obtain
\begin{equation*}
\frac{1}{n} f_0^{\top} (I - S)^2 f_0 \geq \frac{n\wb{r}^{d + 2}}{4\lambda_1} \geq c.
\end{equation*}

\paragraph{Proof of Lemma~\ref{lem:lb_variance}.}
The inequality~\eqref{eqn:lb_variance_2} is immediate.

To prove the inequality~\eqref{eqn:lb_variance_1}, we note that by assumption
\begin{equation*}
1 + \lambda_j\rho \leq 
\begin{cases*}
2,& ~~\textrm{for all $j < k$} \\
2\lambda_j\rho, & ~~\textrm{for all $j \geq k$.}
\end{cases*}
\end{equation*}
and therefore
\begin{equation}
\label{pf:lb_variance_1}
\mathbb{V}(\wh{f}) = \frac{1}{n}\sum_{j = 1}^{n} \frac{1}{(\rho \lambda_j + 1)^2} \geq \frac{1}{4n} \biggl( k - 1 + \sum_{j = k}^{n} \frac{1}{\rho^2 \lambda_j^2} \biggr)
\end{equation}
To lower bound the tail sum, we rely on the fact that for all $j = 1,\ldots,n$,
\begin{equation*}
\lambda_j \leq C j^{-2/d} n\wb{r}^{d + 2}
\end{equation*}
and so $\rho \lambda_j \leq C j^{-2/d} \wb{\rho}$. Therefore,
\begin{align*}
\sum_{j = k}^{n} \frac{1}{\rho^2 \lambda_j^2} & \geq \frac{c}{\wb{\rho}^2} \sum_{j = k}^{n} j^{-4/d} \\
& \geq  \frac{c}{\wb{\rho}^2} \int_{k + 1}^{n} x^{-4/d} \,dx \\
& = \frac{c}{\wb{\rho}^2}\biggl[n^{1 - 4/d} - (k + 1)^{4/d - 1}\biggr] \\
& \geq \frac{c}{\wb{\rho}^2}n^{1 - 4/d}
\end{align*}
and by plugging back in to~\eqref{pf:lb_variance_1} we obtain~\eqref{eqn:lb_variance_1}.

\end{document}

### 2021-02-02 notes/manuscripts/AISTATS/archive/aistats_just_in_case_4.tex
Tidying up.

--- 2020-10-14 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing4.tex
@@
 \DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
+\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
+\DeclareMathAccent{\wb}{0}{mathx}{"73}
+
+\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
+\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}
+
+% Make a widecheck symbol (thanks, Stack Exchange!)
+\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
+\DeclareFontShape{U}{mathx}{m}{n}{
+	<5> <6> <7> <8> <9> <10>
+	<10.95> <12> <14.4> <17.28> <20.74> <24.88>
+	mathx10
+}{}
 \DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
@@
 Additionally, we prove that Laplacian smoothing is manifold adaptive: if $\Xset$ is an $m$ dimensional manifold embedded in $\Reals^d$ with $m \ll d$, then the error rate of Laplacian smoothing methods depends only on the intrinsic dimension $m$ and not on the ambient dimension $d$.
-Empirically, we show that Laplacian smoothing appears not to be minimax optimal when $d > 4$.
+Empirically, we show that Laplacian smoothing appears not to be minimax optimal over $H^1(\Xset)$ when $d > 4$.
 \end{abstract}
@@
 \section{INTRODUCTION}
-
-Suppose we observe responses according to the signal plus noise model:
+In the nonparametric regression problem, we observe samples $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to the signal plus noise model
 \begin{equation}
 \label{eqn:signal_plus_noise_model}
-Y_i = f_{0,i} + \varepsilon_i
+Y_i = f_{0}(X_i) + \varepsilon_i,
 \end{equation}
@@
 
-The Laplacian smoothing estimator $\wh{f}$ \citep{smola2003} is a penalized least squares estimator, defined on a graph.
-Letting $G = \bigl(\{1,\ldots,n\},W\bigr)$ be a weighted graph over the vertices $\{1,\ldots,n\}$,
+The Laplacian smoothing estimator $\wh{f}$ \citep{smola2003} is a penalized least squares estimator.
+It is defined with respect to the weighted graph $G = \bigl(\{1,\ldots,n\},W\bigr)$ over the vertices $\{1,\ldots,n\}$ (equivalently, over the design points $X_1,\ldots,X_n$) as
 \begin{equation}
 \label{eqn:laplacian_smoothing}
-\wh{f} :=  \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap f \biggr\},
+\wh{f} :=  \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap f \biggr\}.
 \end{equation}
-where $\Lap$ is the graph Laplacian matrix (defined in Section 2), and the penalty
+Here $\Lap$ is the graph Laplacian matrix (defined in Section~\ref{sec:problem_setup_and_background}), and the penalty
 \begin{equation*}
@@
 \label{eqn:laplacian_smoothing_test}
-\wh{T} = \frac{1}{n}\bigl\|\wh{f}\bigr\|_2^2 
+\wh{T} := \frac{1}{n}\bigl\|\wh{f}\bigr\|_2^2 
 \end{equation}
@@
 
-Of course there exist many methods for nonparametric regression, but Laplacian smoothing has its own set of advantages. For instance:
+Of course there exist many methods for nonparametric regression \citep{gyorfi2006,wasserman2006,tsybakov2008_book}, but Laplacian smoothing has its own set of advantages. For instance:
 \begin{itemize}
@@
 	\item \emph{Generality.} Laplacian smoothing is well defined whenever one can associate a graph with some observed responses.
-This generality lends itself to many different data modalities: for instance text and image classification, to name only two \citep{kondor2002, belkin03a,belkin2006} \textcolor{red}{(TODO)}: Add a bunch more references here.
+This generality lends itself to many different data modalities: for instance text and image classification, to name only two \citep{kondor2002, belkin03a,belkin2006}.
 	\item \emph{Weak supervision.} Although we study Laplacian smoothing in the supervised learning setup~\eqref{eqn:signal_plus_noise_model}, it can be easily adapted to the semi-supervised or unsupervised settings \citep{nadler09,slepcev17,dunlop2020,calder2019b}.
@@
 \begin{itemize}
-	\item In the \emph{discrete (fixed design) perspective}, both $G$ and $f_0 \in \Reals^n$ are treated as fixed, and the regression ``function'' is really a vector.
+	\item In the \emph{discrete (fixed design) perspective}, the design points $\{X_1,\ldots,X_n\}$ and the graph $G$ are treated as fixed, and inference is carried out on the evaluations $\bigl(f_0(X_1),\ldots,f_0(X_n)\bigr) \in \Reals^n$.
 In this context, tight upper bounds have been derived on the error of graph-based estimators \citep{wang2016, sadhanala16,sadhanala17,kirichenko2017,kirichenko2018} and tests \citep{sharpnack2013,sharpnack2013b,sharpnack2015}, which certify that such procedures are \emph{optimal} over ``function'' classes $\mc{F} \subseteq \Reals^n$.
-The downside of this work is that, depending on the context, it may be unnatural to assume a priori that the graph $G$ is fixed, or that the signal $f_0$ belongs to some discrete ``function'' class.
-	\item In the arguably more natural \emph{continuum (random design) perspective}, in addition to the noisy responses $Y_i$ one observes randomly sampled design points $\{X_1,\ldots,X_n\}$ belonging to a domain $\Xset \subseteq \Rd$, and assumes that $f_{0,i} \equiv f_0(X_i)$ are $n$ evaluations of a regression function $f_0: \Xset \to \Reals$.
-The user then builds a neighborhood graph over the design points, so that $W_{ij}$ is large iff $X_i$ and $X_j$ are close, say in Euclidean distance.
+The downside of this work is that, depending on the context, it may be unnatural to assume a priori that the graph $G$ is fixed, or that the evaluations $\bigl(f_0(X_1),\ldots,f_0(X_n)\bigr)$ belongs to some discrete ``function'' class.
+	\item In the arguably more natural \emph{continuum (random design) perspective},  one assumes the design points $\{X_1,\ldots,X_n\}$ are sampled independently from some distribution $P$ supported on a domain $\Xset$.
+Inference is drawn on the regression function $f_0: \Xset \to \Reals$, which is assumed to be smooth in some sense, e.g.
+it possesses a first derivative bounded in $L^{\infty}$ (Holder) or $\Leb^2$ (Sobolev) norm.
+
+	To conduct graph-based inference on $f_0$, the user builds a neighborhood graph over the design points, so that $W_{ij}$ is large iff $X_i$ and $X_j$ are close, say in Euclidean distance, and then solves e.g.~\eqref{eqn:laplacian_smoothing} or~\eqref{eqn:laplacian_smoothing_test}.
 In this context, various graph-based procedures have been shown to be \emph{consistent}: as $n \to \infty$, they converge to a \emph{continuum} limit (see \citep{belkin07,trillos2018,vonluxburg2008} among others).
-However, until recently such statements were not accompanied by error rates, and such error rates as have been proved \citep{lee2016,trillos2020} are not optimal.
+However, until recently such statements were not accompanied by error rates, and such error rates as have been proved \citep{lee2016,trillos2020} are not optimal over continuum function spaces, such as H\"{o}lder or Sobolev classes.
 \end{itemize}
-Viewed as a whole, this body of work leaves a key question unanswered: what are the optimality properties of graph Laplacian smoothing, viewed as an estimator of a (continuum) regression function?
-This is no small point---rather, it is arguably the fundamental question of nonparametric regression, and without an answer one cannot fully compare the statistical properties of Laplacian smoothing to its competitors.
+Viewed as a whole, this body of work leaves a key question unanswered:
+
+\begin{quote}{What are the optimality properties of graph Laplacian smoothing, viewed as an estimator of a function $f_0$ that is smooth in a continuum sense?} 
+\end{quote}
+On the one hand, this is no small question---rather, it is arguably the fundamental question of nonparametric regression, and without an answer one cannot fully compare the statistical properties of Laplacian smoothing to its competitors.
+On the other hand, it seems difficult to answer: there is a fundamental gap between the \emph{discrete} smoothness imposed by the penalty $f^T \Lap f$ in~\eqref{eqn:laplacian_smoothing} and the \emph{continuum} smoothness assumed on $f_0$, and to obtain sharp upper bounds we needs to bridge this gap in a suitably tight sense.
+
+\section{SUMMARY OF RESULTS}
+In light of this last point, it is worth stepping back for a moment and asking: do we really need to a solve a discrete problem at all?
+After all, we could have instead directly leveraged the (continuum) smoothness of $f_0$ by solving the following variational problem:
+\begin{equation}
+\label{eqn:thin_plate_spline}
+\wt{f} := \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \cdot \int_{\Xset} \|\nabla f(x)\|^2 \,dx \biggr\}
+\end{equation}
+
+\begin{itemize}
+	\item 
+	Of course, there are computational advantages to Laplacian smoothing, as mentioned previously, but let's leave those aside for the second, and ask only whether there is any statistical advantage.
+	\item At first blush, the more likely answer is no.
+As we detail in Section~\ref{sec:minimax_optimal_laplacian_smoothing}, the Laplacian smoothing problem~\eqref{eqn:laplacian_smoothing} and variational problem~\eqref{eqn:thin_plate_spline} are in fact intimately connected; indeed~\eqref{eqn:laplacian_smoothing} can be viewed as a discrete and noisy approximation to~\eqref{eqn:thin_plate_spline}.
+At first blush, this suggests that Laplacian smoothing should at best of inherent the statistical properties thin-plate splines, and at worst possibly have meaningfully larger error.
+	\item However the actual story is quite different: remarkably, Laplacian smoothing outperforms smoothing splines! 
+	\item Laplacian smoothing also demonstrates manifold 
+\end{itemize}
+
+\paragraph{Future extensions.}
 
@@
 \caption{Summary of testing rates over first-order Sobolev balls; rates in black are new results of this paper.
-When $\Xset$ is an $m$ dimensional manifold embedded in $\Rd$, all rates hold with $d$ replaced by $m$.
-Rates for $d \geq 4$ assume $f_0 \in \Leb^4(\Xset,M)$.}
+Rates for $d \geq 4$ assume $f_0 \in \Leb^4(\Xset,M)$.
+When $\Xset$ is an $m$ dimensional manifold embedded in $\Rd$, all rates hold with $d$ replaced by $m$.}
 \label{tbl:testing_rates}
@@
 To avoid any confusion: henceforth when speak of $\wh{f}$ and $\wh{T}$, we will always assume they are fit using the neighborhood graph Laplacian $L = L_{n,r}$.
-
-\paragraph{Sobolev spaces.}
-Now that we've defined the neighborhood graph Laplacian $\Lap_{n,r}$, we step away from graph-based methods for a moment, to briefly recall some classical results regarding minimax rates over Sobolev spaces.
-
-We say that a function $f \in \Leb^2(\Xset)$ belongs to the \emph{first-order Sobolev space} $H^1(\Xset)$ iff for each $j = 1,\ldots,d$, the weak partial derivative $D^jf$ exists and belong to $\Leb^2(\Xset)$.
-For such functions $f \in H^1(\Xset)$, the Sobolev seminorm $\seminorm{f}_{H^{1}(\Xset)}$ is the average size of the gradient;
-\begin{equation*}
-\seminorm{f}_{H^1(\Xset)}^2 = \int_{\Xset} \bigl\|\nabla f(x)\bigr\|^2 \,dx
-\end{equation*}
-with corresponding Sobolev norm $\norm{f}_{H^1(\Xset)} := \norm{f}_{\Leb^2(\Xset)} + |f|_{H^1(\Xset)}$. The Sobolev balls $H^1(\Xset,M)$ for $M > 0$ are then defined as
-\begin{equation*}
-H^1(\Xset,M) := \Bigl\{f \in H^1(\Xset): \norm{f}_{H^1(\Xset)}^2 \leq M^2\Bigr\}.
-\end{equation*}
-For further details regarding Sobolev spaces see e.g \citep{evans10, leoni2017}.
 
@@
 Note that as an alternative to Laplacian smoothing --- which encourages $\wh{f}_1,\ldots,\wh{f}_n$ to satisfy a notion of smoothness over $G_{n,r}$--- we could have directly asked for an estimate $\wt{f}$ which was smooth in the Sobolev sense,
-\begin{equation}
-\label{eqn:thin_plate_spline}
-\wt{f} := \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \cdot \seminorm{f}_{H^{1}(\Xset)}^2\biggr\}
-\end{equation}
+
 where the minimum is taken over all $f \in H^1(\Xset)$.

### 2021-02-11 notes/reference_notes/insights_and_connections.tex
Added updated notes on extension of Laplacian-based estimators.

\documentclass{article}
\usepackage{amsmath, amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{verbatim}
\usepackage{fullpage}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}

% Make a widecheck symbol (thanks, Stack Exchange!)
\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{
	<5> <6> <7> <8> <9> <10>
	<10.95> <12> <14.4> <17.28> <20.74> <24.88>
	mathx10
}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareFontSubstitution{U}{mathx}{m}{n}
\DeclareMathAccent{\widecheck}{0}{mathx}{"71}
% widecheck made

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\iid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}
\newcommand{\lambdavec}{\boldsymbol{\lambda}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}
\newcommand{\Lap}{L}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}
\newcommand{\Id}{I}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Vset}{\mathcal{V}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\SM}{\mathrm{SM}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Theorem environments
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Note on: Insights, Intuition, and Connections regarding Laplacian Learning}
\author{Alden Green}
\date{\today}
\maketitle

We use \emph{Laplacian learning} as a catch-all term for learning algorithms which leverage (possibly discrete approximations of) the Laplace operator to conduct \{unsupervised, semi-supervised, supervised\} learning tasks. In this note, I will detail various physical analogies and intuitions to explain what Laplacian learning is ``doing'', draw connections between Laplacian learning and other types of learning (e.g. kernel smoothing, RKHS regression), and hopefully provide some insight regarding when and why Laplacian learning is a reasonable idea. Much of what follows will be review of others' ideas, and I provide references where appropriate.

\section{Physical analogies and intuitions} 
One theme of the literature regarding Laplacian learning is that algorithms which use the Laplace operator are sometimes well-posed, and sometimes not. We will be particularly interested in analogies and intuition that explain when and why algorithms that use the Laplace operator succeed, and when and why they fail. The question of well-posedness often hinges on the extent to which constraints enforced on some subset of an ambient domain $\Omega$ ensure a well-posed problem: that is, a problem with a reasonable solution on $\Omega$. We will call the constraints \emph{boundary conditions}.

Suppose we observe data $(x_1,y_1),\ldots,(x_n,y_n)$ and additionally $x_{n + 1},\ldots,x_{N}$, with $\Omega_N = \{x_1,\ldots,x_N\} \subset \Rd$. For a given kernel $\eta(\cdot): [0,\infty) \to [0,\infty)$, and length scale (or bandwidth) $\varepsilon$, the Laplace operator $L_{N,\varepsilon}: H^2(\Omega_N) \to L^2(\Omega_N)$ is defined according to, 
\begin{equation*}
L_{N,\varepsilon}u(x) = \frac{1}{N\epsilon^{d + 2}}\sum_{j = 1}^{N} \bigl(u(x) - u(x_j)\bigr)\eta\biggl(\frac{|x - x_i|}{\varepsilon}\biggr).
\end{equation*}
The operator $L_{N,\varepsilon}$ is associated with a \emph{Dirichlet energy} $b_{N,\varepsilon}: H^1(\Omega_N) \to [0,\infty)$, defined as 
\begin{equation*}
b_{N,\varepsilon}(u) = \sum_{i,j = 1}^{N} \bigl(u(x_i) - u(x_j)\bigr)^2 \eta\biggl(\frac{|x - x_i|}{\varepsilon}\biggr),
\end{equation*}
the connection being that $b_{N,\varepsilon}(u) = \dotp{u}{L_{N,\varepsilon}u}_{N}$ for any $u \in H^2(\Omega_N)$. Our goal is to learn a function $u$ on $x_{n + 1},\ldots,x_{N}$ using the labels $Y = \{y_1,\ldots,y_n\}$ and the Laplace operator $L_{N,\varepsilon}$.

\paragraph{Mininum energy interpolation, and a connection with random walks.}

A natural Laplace learning algorithm for semi-supervised learning is the minimum (Dirichlet) energy interpolant of $y_1,\ldots,y_n$, originally considered in the semi-supervised learning literature by \textcolor{red}{Zhu}, (and in the supervised learning case by \textcolor{red}{Duchon}) and defined as 
\begin{equation}
\label{eqn:min_energy_interpolant}
\begin{aligned}
& \wh{u} := \min \|b_r(u)\|^2, \\
~~\textrm{such that}~~ &\wh{u}(x_i) = y_i ~\textrm{for}~ i = 1,\ldots,n.
\end{aligned}
\end{equation}
Some basic algebra shows that ${u}$ satisfies the \emph{Laplace equation with Dirichlet boundary conditions}, i.e. 
\begin{equation}
\label{eqn:laplace_equation}
\begin{cases}
L_{N,\varepsilon}{\wh{u}}(x_i) = 0,& ~~\textrm{$x_i \in \Omega_N \setminus \{x_1,\ldots,x_n\}$} \\
\wh{u}(x_i) = y_i,&~~\textrm{$x_i \in \{x_1,\ldots,x_n\}$.}
\end{cases}
\end{equation}
and takes the explicit form $\wh{u} = L^{\dagger} A y$. Here $A \in \Reals^{N \times n}$ is the kernel smoothing matrix with entries $A_{ij} = \eta(\|x_i - x_j\| / \varepsilon)$.

The solution to the Laplace equation admits a probabilistic interpretation in terms of random walks. Let $X_0,X_1,\ldots$ be a random walk over $x_1,\ldots,x_N$, where at each step the walk moves from its current position $x_i$ to a neighboring point $x_j$ with probability $p(x_i,x_j) = A_{ij}/d(x_i)$, where $d(x_i) = \sum_{j = 1}^{n} A_{ij}$. The generator of this walk is the random walk Laplacian $L_{\mathrm{rw}} := (1/d) L_{N,\varepsilon}$. Suppose we stop this walk the first time it reaches a labeled point, i.e let the stopping time $\tau := \inf\{t: X_t \in \{x_1,\ldots,x_n\}\}$. Denote the resulting label by $y_{\tau}$. The solution $\wh{u}$ to~\eqref{eqn:min_energy_interpolant} is then the expected value of the response when the random walk stops,
\begin{equation}
\label{eqn:random_walk_interpretation_laplace_learning}
\wh{u}(x) := \Ebb[y_{\tau}|X_0 = x].
\end{equation}

\paragraph{Resistance distance, a connection with random walks, and two connections with Laplacian learning.}
Another way we can use the Laplacian is to define a metric between points $x_i$ and $x_j$. The metric we will use is the \emph{resistance distance}, defined as 
\begin{equation*}
R_{ij} = R(x_i,x_j) = \bigl \langle L_{N,\varepsilon}^{\dagger}(e_i - e_j),(e_i - e_j)\bigr \rangle_N
\end{equation*}
The resistance distance arises from two different (and therefore equivalent) physical interpretations. The first treats the graph $G$ corresponding to adjacency matrix $A$ as a resistor network. Placing an external current $c_{ext} := e_i - e_j$ then induces a potential difference between points $x_i$ and $x_j$, and it follows from Ohm's Law and Kirchoff's Laws that this potential difference is precisely $R(i,j)$. The second interpretation concerns the random walk $X_0,X_1,\ldots$ with starting point $X_0 = x_i$. The resistance distance $R_{ij}$ is (up to a rescaling) the \emph{commute time} of this random walk. The commute time $C_{ij} = C(x_i,x_j)$ is the sum  of two \emph{hitting times}
---$H(x_i,x_j)$, the expected time before the walk hits point $x_j$, plus $H(x_j,x_i)$, the expected time it takes for the random walk to return to point $x_i$.

The first connection between resistance distance and Laplace learning is simply that both can be interpreted in terms of these hitting times. We have just seen that the effective resistance between points $x_i$ and $x_j$ can be written as $R(i,j) \propto H(i,j) + H(j,i)$. On the other hand,~\eqref{eqn:min_energy_interpolant} provides a way to relate $\wh{u}$ to the expected hitting time. Concretely, supposing at the labeled points $y_i = g(x_i)$ for some Lipschitz function $g$ with Lipschitz constant $L$, then at an unlabeled point $x \in \Omega_N$ a martingale argument shows that
\begin{equation*}
\bigl|\wh{u}(x) - g(x)\bigr|^2 \lesssim \Ebb[\tau] \cdot \Ebb[|X_{\tau} - x|^2]  \cdot L,
\end{equation*}
where $\Ebb[\tau] = H\bigl(x,\{x_1,\ldots,x_n\}\bigr)$. \textcolor{red}{TODO: Calder2020a makes a lower bound argument, which also belongs here.}

The second connection between effective resistance and Laplace learning is an exact coupling between the two. Suppose we solve the minimum energy interpolation problem with $n = 2$, $y_1 = 1$ and $y_2 = -1$. Then the effective resistance is exactly one over the energy of $\wh{u}$,
\begin{equation*}
b_{\varepsilon}(\wh{u}) \cdot R(1,2) = 1.
\end{equation*}

\paragraph{Connections between resistance distance and Laplace learning: ill-posed regime.}
We know that, when dimension $d \geq 2$, both the resistance distance and the minimum-energy interpolant are ill-posed, meaning they converge to degenerate solutions as the number of unlabeled points $N \to \infty$, and the radius $\varepsilon \to 0$. The resistance distance, which for finite $N$ is a measure of some mixture of global and local information, in the limit contains only global information:
\begin{equation}
\label{eqn:resistance_distance_illposed}
\lim_{\substack{N \to \infty \\ \varepsilon \to 0}}H(x_i,x_j) \propto \frac{1}{p(x_j)} \Longrightarrow \lim_{\substack{N \to \infty \\ \varepsilon \to 0}} R(x_i,x_j) \propto \frac{1}{p(x_i)} + \frac{1}{p(x_j)}.
\end{equation}
The minimum-energy interpolant $\wh{u}$, which for finite $N$ propagates the labels $Y$ in some way that weights both local and global information, in the limit interpolates $Y$ with a constant trend function: for \emph{a constant value of $n$}, 
\begin{equation}
\label{eqn:min_energy_interpolant_illposed}
\lim_{\substack{N \to \infty \\ \varepsilon \to 0}} \wh{u}(x_i) = 
\begin{cases*}
\wb{y},& ~~\textrm{if $x_i \in \Omega_N\setminus\{x_1,\ldots,x_n\}$},\\
y_i,& ~~\textrm{if $x_i \in \{x_1,\ldots,x_n\}$}
\end{cases*}
\end{equation}

A natural question: can we use the connections of the previous section to relate the ill-posedness of Laplacian learning and resistance distance? The answer is yes. The second connection shows \emph{that} the ill-posedness of Laplacian learning implies the ill-posedness of resistance distance, and vice versa. The first connection allows us to use the ill-posedness of resistance distance to shed light on \emph{why} Laplace learning is ill-posed. 
\begin{itemize}
	\item \textbf{Second connection.} The convergence of the resistance distance in~\eqref{eqn:resistance_distance_illposed} implies that a constant function plus spikes is a minimum-energy interpolant, i.e. that $\wh{u}$ must satisfy~\eqref{eqn:min_energy_interpolant_illposed}. Conversely, the structure of the minimum-energy interpolant as stated in~\eqref{eqn:min_energy_interpolant_illposed} implies that the resistance distance must converge to the degenerate limit~\eqref{eqn:resistance_distance_illposed}.
	\item \textbf{First connection.} The random walk has mixed before it has a probability of hitting any labeled node.
	\item Explain why increasing the radius actually exacerbates the problem---it simply decreases the mixing time. In fact, as \textcolor{red}{(Nadler09)} point out, if we take $N \to \infty$ and hold $\varepsilon$ constant, then estimator $\wh{u}$ will be degenerate even when $d = 1$.
\end{itemize}

\paragraph{Connections between resistance distance and Laplace learning: well-posed regime.}

\begin{itemize}
	\item Explain what the well-posed regime is: when limits are taken simultaneously as $n,N \to \infty$, with $n$ satisfying a certain scaling $n = a(N)$. For instance, the linear scaling $n = \Theta(N)$. 
	\item Consider a kernel smoother which makes explicit use of resistance distance, e.g.
	\begin{equation*}
	\wt{u}(x_i) = \frac{\sum_{j = 1}^{n} K(R(i,j)) u(x_j)}{\sum_{j = 1}^{n} K(R(i,j))}
	\end{equation*}
	for some kernel function $K(\cdot)$.

	Make the banal mathematical point that even if $R(x_i,x_j)$ are all within $\epsilon$ of $1/d(x_j) + 1/d(x_i)$, if there are an increasing number of labeled points, this does not imply anything about the distance between $\wt{u}$ and $\wb{y}$.
	\item Conclude that the crucial point is the ratio of labeled to unlabeled points, as this determines whether the walk will have mixed before it arrives at a labeled point.
	\item Discuss other fixes---three of which are ``smearing the labels''---I have not seen anyone mention this, to the best of my knowledge---changing the graph \textcolor{red}{(Shi, Calder)}, and changing the norm \textcolor{red}{(El-Alaoui, Thorpe)}. 
\end{itemize}

\paragraph{Future directions.}

To be clear, by future directions I mean only things for me to track down in the future, rather than any claim of novelty. A tremendous amount of work has been done regarding Laplace learning and I have certainly only scratched the surface of it in this summary.
\begin{itemize}
	\item In practice, people often solve the penalized version of~\eqref{eqn:min_energy_interpolant}. This is thought to better handle the common situation of noisy responses. We refer to this as Laplacian smoothing. It would be interesting to find similar insights regarding Laplacian smoothing.
	\item There are other relevant diffusion-based perspectives on Laplacian learning, i.e. tying Laplacian learning to walks or diffusions over graphs. This notably includes the \emph{diffusion maps} perspective of \textcolor{red}{Coifman and Lafon}. These perspectives should be reviewed as well. 
	\item As indicated by the discussion in the ill-posed section, the failure of Laplace learning when $d \geq 2$ is due to the fact that the walk $X_0,X_1,\ldots$ mixes before a labeled node is ever reached. A natural fix, then, is to simply prevent the random walk from mixing. For instance, we could pick a threshold $T$; if after $T$ steps the random walk has not arrived at a labeled point, restart. Alternatively, we might consider the random walk $\wt{X}_0,\wt{X}_1,\ldots$, which at each step has probability $(1 - \alpha)$ of moving to an adjacent point, and probability $\alpha$ of returning to the seed node $\wt{X}_0$. In other words, run PPR.
	\item It has been observed \textcolor{red}{(Nadler09, Calder2020)} that even in the ill-posed regime---concretely, when $N$ is finite but extremely large compared to $n$---although $\wh{u}$ is quite obviously a bad estimator, once appropriately re-centered it performs quite well at classification tasks. This phenomenon has not been theoretically explained. 
\end{itemize}

\section{Laplacian learning as Reproducing Kernel Hilbert Space regression}
We refer to the penalized analog of~\eqref{eqn:min_energy_interpolant} as \emph{Laplacian smoothing}, defined for a given value of $\lambda$ as
\begin{equation*}
\wt{u} := \argmin \|y - u\|_n^2 + \lambda b_r(u)
\end{equation*}
When the response is zero-mean, Laplacian smoothing is equivalent to ridge regression in the Reproducing Kernel Hilbert Space (RKHS) defined by the kernel $K = L^{\dagger}$. It is immediate that the kernel $K$ defines an RKHS $H_0^1(\Omega_N)$: $L^{\dagger}$ is a positive semi-definite (psd) bilinear form, and it is automatically trace-class because $L^2(\Omega_N)$ is finite-dimensional. The feature map is $\phi: x \mapsto \bigl(\lambda_2^{1/2} v_2(x),\lambda_3^{1/2} v_3(x),\ldots,\lambda_N^{1/2} v_N(x)\bigr)$. The norm is $\|u\|_{H_0^1(\Omega_N)} = \sqrt{b_{\varepsilon}(u)}$.

\paragraph{Laplacian smoothing as RKHS regression: ill-posed regime.}
When $d \geq 2$, as $N \to \infty$ and $\varepsilon \to 0$ (second condition optional), and $\lambda$ remains fixed, $\wt{u}$ will converge to the ill-posed limit~\eqref{eqn:min_energy_interpolant_illposed}. This can be explained by the fact that while for any finite $N$, $L^{\dagger}$ is trace-class, when $d \geq 2$ asymptotically the trace explodes, meaning
\begin{equation*}
\lim_{\substack{N \to \infty \\ \varepsilon \to 0}} tr(L^{\dagger}) \to \infty.
\end{equation*}
This will not be true when $d = 1$.

Equivalently, while it is the case that the evaluation functional $L_x: f \mapsto f(x)$ is bounded for any finite $N$,
\begin{equation*}
|L_x(f)| \leq C(N) \cdot \|f\|_{H_0^1(\Omega_N)},
\end{equation*}
asymptotically the statement is not true: $\lim_{N \to \infty} C(N) = \infty$.

\paragraph{Laplacian learning and RKHS regression: well-posed regime.}
We now give a simple example showing that when $n/N \gg \varepsilon^2$, the representation cost (in $H_0^1(\Omega_N)$) of a spiky interpolant such as~\eqref{eqn:min_energy_interpolant_illposed} explodes. This is known to be exactly the minimum labeling rate at which minimum Dirichlet energy interpolation converges to a sensible continuum limit.

Suppose $y_i = g(x_i)$ for some function $g$, and suppose (for convenience only) that $\wb{y} = 0$. Then $u_N = (y_1,\ldots,y_n, 0,\ldots,0)$ belongs to $H_0^1(\Omega_n)$ and has squared-norm $\|u_N\|_{H_0^1(\Omega_N)}^2 = b_r(u_N)$ of at least
\begin{equation*}
b_{\varepsilon}(u_N) \geq \frac{1}{N^2 \varepsilon^{d + 2}} \sum_{i = 1}^{n} \sum_{j = n + 1}^{N} g(x_i)^2 \geq \frac{n}{N^2 \varepsilon^{d + 2}} d_{\min}(G) \|g\|_n^2.
\end{equation*} 
In the above $d_{\min}(G) = \min\bigl\{\sum_{j = n + 1}^{N} A_{ij}: i = 1,\ldots,n\bigr\}$. Under standard \{random, fixed\} design conditions on $\Omega_N$, asymptotically $d_{\min}(G) \asymp N\varepsilon^d$ and $\|g\|_n^2 \to \|g\|_{\Leb^2(\Omega)} > 0$. Thus
\begin{equation*}
\lim_{N \to \infty} b_{\varepsilon}(u_N) \geq c \cdot \lim_{N \to \infty} \frac{n}{N \varepsilon^2}.
\end{equation*}
This is to be contrasted with the interpolator $u = \bigl(g(x_1),\ldots,g(x_n), g(x_{n + 1}),\ldots,g(x_N)\bigr)$, which has finite representational cost in the limit as $n \to \infty, N \to \infty$, regardless of the scaling of $n$ with respect to $N$.

\paragraph{Future directions.}
\begin{itemize}
	\item Relate the minimum Dirichlet interpolant to overparameterized minimum-norm linear regression. 
\end{itemize}

\section{Laplacian learning as kernel smoothing}

\end{document}

### 2021-02-11 notes/weekly_notes/12.24.20.NotesForWeek.tex
Added updated notes on extension of Laplacian-based estimators.

\documentclass{article}
\usepackage{amsmath, amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=blue,
	citecolor=blue
}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{verbatim}
\usepackage{fullpage}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}

% Make a widecheck symbol (thanks, Stack Exchange!)
\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{
	<5> <6> <7> <8> <9> <10>
	<10.95> <12> <14.4> <17.28> <20.74> <24.88>
	mathx10
}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareFontSubstitution{U}{mathx}{m}{n}
\DeclareMathAccent{\widecheck}{0}{mathx}{"71}
% widecheck made

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\iid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}
\newcommand{\lambdavec}{\boldsymbol{\lambda}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}
\newcommand{\Lap}{L}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}
\newcommand{\Id}{I}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Vset}{\mathcal{V}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\SM}{\mathrm{SM}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Theorem environments
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Note on: Out-of-Sample Error for Laplacian Eigenmaps}
\author{Alden Green}
\date{\today}
\maketitle

The graph Laplacian eigenmaps estimator $\wh{\theta}_{\LE}$ is a least-squares estimator, where a response vector $Y = (Y_1,\ldots,Y_n)$ is projected onto the span of the first $K$ eigenvectors of a graph Laplacian matrix. It is defined only in-sample, whereas out-of-sample error is more interesting in many applications. There is no commonly agreed upon way in the literature to extend $\wh{\theta}_{\LE}$, or indeed any type of Laplacian estimator. In this note, I discuss three methods for doing so: least-squares regression of $\wh{\theta}_{\LE}$ onto an a priori defined set of features, Nadaraya-Watson kernel smoothing, treating $\wh{\theta}_{\LE}$ as the response, and Nystr\"{o}m extension of Laplacian eigenvectors. These methods are roughly ordered from least to most preferable, but also (as these things go) easiest to hardest to analyze. I begin in Section~\ref{sec:least_squares_arbitrary_design_and_features} by reviewing known bounds on the out-of-sample error of a (generic) least-squares estimator; these bounds will be useful for the analysis of both the first and third methods.

\section{Out-of-sample error for least-squares estimators: arbitrary design and features}
\label{sec:least_squares_arbitrary_design_and_features}
This section largely summarizes results from a paper by Daniel Hsu on random design regression. We start with some general setup that will apply to everything that follows.

Suppose we observe $(X_1,\wh{\theta}_1),\ldots,(X_n,\wh{\theta}_n)$, where $X_1,\ldots,X_n \in \Reals^d$ and $\wh{\theta}_{1:n} = (\wh{\theta}_1,\ldots,\wh{\theta}_n) \in \Reals^n$. Our goal is to use this observed data to construct an estimator $\wh{f}$ which is on average close to $f^{\ast}: \Omega \to \Reals$ on the domain $\Omega$ (for instance, $\Omega = [0,1]^d$.) Specifically, we will consider the following least-squares estimator,
\begin{equation*}
\wh{\beta} = \argmin \| \wh{\theta}_{1:n} - \Phi_n^{\top} \beta \|_n^2 = (\Phi_n^\top \Phi_n)^{-1}\Phi_n^{\top} \wh{\theta}_{1:n},~~ \wh{f} = \sum_{k = 1}^{\kappa} \wh{\beta}_k \varphi_k
\end{equation*}
where $\varphi_1,\ldots,\varphi_{K}$ are each elements of $\Leb^2(\Omega)$, and $\Phi_n \in \Reals^{n \times K}$ has entries $(\Phi_{n})_{ik} = \varphi_k(X_i)$. If $\wh{\theta}_{1:n}$ lies in the span of the columns of $\Phi_n$, then $\wh{f}$ is a interpolant of $\wh{\theta}_{1:n}$; this applies to our third method, Nystr\"{o}m extension of Laplacian eigenvectors, but not to our first method. We will measure error in the $\Leb^2(P)$ norm $\norm{\wh{f} - f^{\ast}}_P^2 = \int_{\Omega} (\wh{f}(x) - f^{\ast}(x))^2 \,dP(x)$, where $P$ is a probability measure supported on domain $\Omega$. For this section, we assume no relationship between $P$ and $X_1,\ldots,X_n$.

We begin by decomposing this norm into several modular terms, which capture the different types of error intrinsic to the problem. To this end, let
\begin{align*}
\wt{\beta} & = \argmin_{\beta \in \Reals^K} \|f^{\ast}_{1:n} - \Phi_n^{\top} \beta \|_n^2 = (\Phi_n^\top \Phi_n)^{-1}\Phi_n^{\top} {f}^{\ast}_{1:n},&& \wt{f}_K = \sum_{k = 1}^{K} \wt{\beta}_k \varphi_k \\ 
\beta^{\star} & = \argmin_{\beta \in \Reals^K} \|f^{\ast} - \sum_{k = 1}^{K} \beta_k \varphi_k \|_P^2,&& f_K^{\ast} = \sum_{k = 1}^{K} \beta_k^{\ast} \varphi_k
\end{align*}
and let $f_{\perp}^{\ast} = f^{\ast} - f_{K}^{\ast}$. We decompose $\norm{\wh{f} - f^{\ast}}_P^2$ as follows:
\begin{equation*}
\norm{\wh{f} - f^{\ast}}_P^2 \leq 3\Bigl(\underbrace{\norm{\wh{f} - \wt{f}_K}_P^2}_{\textrm{estimation error}} + \underbrace{\norm{\wt{f}_K - f_K^{\ast}}_P^2}_{\textrm{sampling error}} + \underbrace{\norm{f_K^{\ast} - f^{\ast}}_P^2}_{\textrm{truncation error}}\Bigr)
\end{equation*}
This leaves us with three separate kinds of error, and we now upper bound the first two.

\paragraph{Notation.} Let $e_{1:n} = \wh{\theta}_{1:n} - {f}^{\ast}_{1:n}$. Let~$\Sigma \in \Reals^{K \times K}$ be the second moment matrix of $\varphi_1,\ldots,\varphi_K$ with $(k,\ell)$-th entry $\Sigma_{k\ell} = \Ebb_{X \sim P}\bigl[\varphi_k(X) \varphi_{\ell}(X)\bigr]$. Likewise let $\wh{\Sigma} = \Phi_n^{\top} \Phi_n/n$ be the empirical second moment matrix. For a matrix $A \in \Reals^{K \times K}$, let $\|A\|$ denote the operator norm of $A$.  For a probability distribution $Q$ on $\Omega$, let $d_{\min}(Q) = \inf_{x \in \Omega} \int K(\|z - x\|/r) \,dQ(z)$ be the minimum degree of $Q$.

\paragraph{Estimation error.}
Noting that
\begin{equation*}
~~\wh{\beta} - \wt{\beta} = \frac{1}{n}\wh{\Sigma}^{-1} \Phi_n^{\top} e_{1:n},
\end{equation*}
the estimation error satisfies the inequality,
\begin{align*}
\norm{\wh{f} - \wt{f}_K}_P^2 & = (\wh{\beta} - \wt{\beta})^{\top} \Sigma (\wh{\beta} - \wt{\beta}) \\
& = \frac{1}{n^2} e_{1:n}^{\top} \Phi_n \wh{\Sigma}^{-1} \Sigma \wh{\Sigma}^{-1} \Phi_n^{\top} e_{1:n} \\
& \leq \|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\| \cdot \|e_{1:n}\|_n^2.
\end{align*}
with the inequality following since $\|AB\| = \|BA\|$.

\paragraph{Sampling error.}
Multiplying $\wt{\beta} - \beta^{\star}$ on the left by $\Sigma^{1/2}$, we have
\begin{equation*}
\Sigma^{1/2}(\wt{\beta} - \beta^{\ast}) =  \Sigma^{1/2} \wh{\Sigma}^{-1} \bigl( \frac{1}{n} \Phi_n^{\top} f_{1:n}^{\ast} - \frac{1}{n} \Phi_n^{\top} \Phi_n \beta^{\ast} \bigr) = \Sigma^{1/2} \wh{\Sigma}^{-1} \frac{1}{n} \Phi_n^{\top}  \bigl(\underbrace{f_{1:n}^{\ast} - \Phi_n \beta^{\ast}}_{ = (f_{\perp}^{\ast})_{1:n}} \bigr) = \frac{1}{n}\Sigma^{1/2} \wh{\Sigma}^{-1} \Phi_n^{\top} (f_{\perp}^{\ast})_{1:n}.
\end{equation*}
Therefore
\begin{equation*}
\|\wt{f}_K - f_K^{\ast}\|_P^2  = \|\wt{\beta} - \beta^{\ast}\|_{\Sigma}^2 \leq \|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\| \cdot \Bigl\|f_{\perp}^{\ast}\|_n^2
\end{equation*}

The results of this section are summarized by Lemma~\ref{lem:l2p_error_fixed_design}. 
\begin{lemma}
	\label{lem:l2p_error_fixed_design}
	Suppose $\wh{\Sigma} \succeq 0$. Then the $\Leb^2(P)$ error $\|\wh{f} - f^{\ast}\|_P^2$ is upper bounded as follows:
	\begin{equation}
	\label{eqn:l2p_error_fixed_design}
	\|\wh{f} - f^{\ast}\|_P^2 \leq 3\Bigl(\|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\| \cdot \|\wh{\theta}_{1:n} - {f}^{\ast}_{1:n}\|_n^2 + \|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\| \cdot \|f_{\perp}^{\ast}\|_n^2 + \|f_{\perp}^{\ast}\|_P^2\Bigr).
	\end{equation}
	If $\wh{\Sigma} = I$ then
	\begin{equation}
	\label{eqn:l2p_error_fixed_design_identity}
	\|\wh{f} - f^{\ast}\|_P^2 \leq 3\Bigl(\|\Sigma\| \cdot \|\wh{\theta}_{1:n} - {f}^{\ast}_{1:n}\|_n^2 + \|\Sigma\| \cdot \|f_{\perp}^{\ast}\|_n^2 + \|f_{\perp}^{\ast}\|_P^2\Bigr).
	\end{equation}
\end{lemma}
(Basically) no assumptions so far have been made on $X_1,\ldots,X_n$ or $\varphi_1,\ldots,\varphi_K$. In the standard setup $X_1,\ldots,X_n$ are randomly sampled from $P$, and $\varphi_1,\ldots,\varphi_K$ are a priori determined. Then matrix Bernstein's inequality gives an upper bound on $\|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\|$, standard Bernstein's inequality gives an upper bound on $\|f_{\perp}^{\ast}\|_n^2 - \|f_{\perp}^{\ast}\|_P^2$, and we obtain from~\eqref{eqn:l2p_error_fixed_design} an upper bound that depends on $\|e_{1:n}\|_n^2$ and $\|f_{\perp}^{\ast}\|_P^2$.

If $\varphi_1,\ldots,\varphi_{K}$ are chosen in a data-dependent manner, then the analysis is plainly quite different. Each of $\|\Sigma\|$, $\|f_{\perp}^{\ast}\|_n^2$ and $\|f_{\perp}^{\ast}\|_P^2$ are random variables which depend on $\varphi_1,\ldots,\varphi_K$. We also do not have concentration of $\|f_{\perp}^{\ast}\|_n^2$ around $\|f_{\perp}^{\ast}\|_P^2$.

\section{Kernel smoothing}
Suppose we observe $(X_1,\wh{\theta}_1),\ldots,(X_n,\wh{\theta}_n)$, where $X_1,\ldots,X_n \in \Reals^d$ are sampled i.i.d from an unknown distribution $P$ on domain $\Omega$, and $\wh{\theta}_{1:n} = (\wh{\theta}_1,\ldots,\wh{\theta}_n) \in \Reals^n$. Our goal is to construct an estimator $\wh{f}$ such that the error $\|\wh{f} - f^{\ast}\|_P^2$ is as small as possible relative to $\|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n^2$.

For a kernel function $\eta(\cdot): [0,\infty) \to (-\infty,+\infty)$, bandwidth $\varepsilon > 0$, and a distribution $Q$, the \emph{Nadaraya-Watson kernel smoother} $T_{\varepsilon,Q}$ is given by
\begin{equation*}
\bigl(T_{\varepsilon,Q}f)(x) := \frac{1}{d_Q(x)} \int_{\Omega} f(z)\eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dQ(z)
\end{equation*}
where $d_Q(x) = \int_{\Omega} \eta\bigl(\|z - x\|/\varepsilon\bigr) \,dQ(z)$. We denote $T_{\varepsilon,P_n}$ by $T_{\varepsilon,n}$, and $d_{P_n}$ by $d_n$. For convenience, we will assume that $\eta$ is normalized, $\int_{0}^{\infty} \eta(\|z\|) \,dz = 1$, and we will denote $K(\cdot) = (\eta(\cdot))^2$.

The estimate we will consider is a kernel smoother passed over $\wh{\theta}$,
\begin{equation*}
\wh{f}(x) := T_{\varepsilon,n}\wh{\theta}_{1:n}(x).
\end{equation*}
In order to relate the out-of-sample error $\|\wh{f} - f^{\ast}\|_P$ to the in-sample error $\|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n$, we will assume the following.
\begin{enumerate}[label=(A\arabic*)]
	\item
	\label{asmp:kernel}
	The kernel function $\eta$ is supported on a subset of $[0,1]$, with $\eta(1/2) > 0$. Additionally, $\eta$ is \textcolor{red}{Lipschitz} on its support.
	\item 
	\label{asmp:domain} There exists $\alpha > 0$ such that for every $x \in \mc{X}$, $d_{\nu}(x) \geq \alpha \cdot \varepsilon^d$.
	\item 
	\label{asmp:density}The distribution $P$ admits a density $p$ with respect to Lebesgue measure. The density $p \in C^1(\mc{X})$, and is bounded away from $0$ and $\infty$ for all $x \in \mc{X}$,
	\begin{equation*}
	0 < p_{\min} \leq p(x) < p_{\max} < \infty.
	\end{equation*}
\end{enumerate}
Under Assumption~\ref{asmp:kernel}, we automatically have that $d_\nu(x) = \varepsilon^d$ for all $x \in \mc{X}_{\varepsilon}$. Assumption~\ref{asmp:domain} is thus a condition on the boundary $\partial \Omega$, and is implied for all sufficiently small $\varepsilon$ if the boundary is Lipschitz. 
\begin{lemma}
	\label{lem:kernel_smoothing_laplacian_eigenmaps}
	Suppose $f^{\ast} \in H^1(\mc{X})$ and $p \in C^1(\mc{X})$. Then with probability at least $1 - \delta - C\varepsilon^d\exp\{-Cn\varepsilon^d\}$, it holds that
	\begin{equation*}
	\|\wh{f} - f^{\ast}\|_P^2 \leq C\biggl(\|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n^2 + \frac{1}{\delta} \cdot \frac{\varepsilon^2}{n\varepsilon^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{\varepsilon,P}f^{\ast} - f^{\ast}\|_P^2\biggr)
	\end{equation*}
\end{lemma}
Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data (on the order of $1/(n\varepsilon^d)$) by a factor of $\varepsilon^2$. The bias term, on the other hand, is standard, and in particular the following estimate of the bias can be obtained. 
\begin{lemma}
	Suppose $f^{\ast} \in H_0^{s}(\Omega)$, $p \in C^{\beta}(\Omega)$ for $\beta = s - 1$, and that $\eta$ is an order-$s$ kernel. Then
	\begin{equation*}
	\|T_{P,\varepsilon}f^{\ast} - f^{\ast}\|_P^2 \leq C \varepsilon^{2s}.
	\end{equation*}
\end{lemma}

Thus balancing bias and variance, using a smaller bandwidth ($\varepsilon \asymp n^{-1/(2\beta + d)}$) than is typical for kernel smoothing with noisy data, leads to an asymptotically negligible error ($n^{-2s/(2\beta + d)}$) compared to the minimax rate for estimating Sobolev functions.

\begin{proof}[Proof of Lemma~\ref{lem:kernel_smoothing_laplacian_eigenmaps}]
	By the triangle inequality, 
	\begin{equation*}
	\|\wh{f} - f^{\ast}\|_P \leq \|\wh{f} - T_{\varepsilon,n}f^{\ast}_{1:n}\|_P + \|T_{\varepsilon,n}f^{\ast}_{1:n} - f^{\ast}\|_P.
	\end{equation*}
	We analyze each of the two terms in the above expression separately. We will show that
	\begin{equation}
	\label{pf:kernel_smoothing_laplacian_eigenmaps_0}
	\|\wh{f} - T_{\varepsilon,n}f^{\ast}_{1:n}\|_P \leq \frac{1}{\alpha}\bigl(1 + o_P(1)\bigr) \cdot \|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n
	\end{equation}
	and 
	\begin{equation}
	\label{pf:kernel_smoothing_laplacian_eigenmaps_0.5}
	\|T_{\varepsilon,n}f^{\ast}_{1:n} - f^{\ast}\|_P^2 \leq \frac{1}{\delta}\frac{\varepsilon^2}{n\varepsilon^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{\varepsilon,P}f^{\ast} - f^{\ast}\|_P^2,
	\end{equation}
	the latter inequality holding with probability $1 - \delta$.

	\textit{Proof of \eqref{pf:kernel_smoothing_laplacian_eigenmaps_0}.} We will begin by assuming that $\eta(z) \geq 0$ for all $z$, and return to prove the claim in its full generality at the end.

	The left hand side of~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is equal to $\|T_{\varepsilon,n}(\wh{\theta}_{1:n} - f^{\ast}_{1:n})\|_P$, and~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is implied by showing that the operator $T_{\varepsilon,n}: L^2(X) \to L^2(\Omega)$ is (up to constant) a contraction, meaning that for any $u \in L^2(X)$,
	\begin{equation}
	\label{pf:kernel_smoothing_laplacian_eigenmaps_1}
	\|T_{\varepsilon,n}u\|_P^2 \leq \frac{1}{\alpha}\bigl(1 + o_P(1)\bigr) \|u\|_n^2.
	\end{equation}
	We begin with a pointwise estimate, which follows from Jensen's inequality:
	\begin{equation*}
	\Bigl(\bigl(T_{n,\varepsilon}u\bigr)(x)\Bigr)^2 \leq \frac{1}{d_n(x) n} \sum_{i = 1}^{n} u_i^2 \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
	\end{equation*}
	(Note that it is here that we have used the positivity of $\eta(\cdot)$.) Integrating with respect to $P$ and exchanging sum with integral yields an upper bound on the $L^2(P)$ norm of $T_{n,\varepsilon}u$,
	\begin{equation}
	\label{pf:kernel_smoothing_laplacian_eigenmaps_2}
	\|T_{n,\varepsilon}u\|_P^2 \leq \sum_{i = 1}^{n} u_i^2 \Bigl(\int_{\Omega} \frac{\eta(\|X_i - x\|/\varepsilon)}{n d_n(x)} p(x) \,dx \Bigr)
	\end{equation}
	The degree estimate $d_n(x)$ is quite close to $p(x)$. As we show in Lemma~\ref{lem:measure_of_ball_estimate}, if $\varepsilon = \omega((\log n/n)^{1/d})$ then the empirical measure $d_n(x)$ concentrates around $d_P(x)$ uniformly over $x \in \Omega$, 
	\begin{equation*}
	\sup_{x \in \Omega}|d_n(x) - d_P(x)| = o_P(\varepsilon^d);
	\end{equation*}
	moreover, it follows from Lipschitz continuity of $p$ that for any $x \in \Omega$,
	\begin{equation*}
	d_P(x) = \int_{\Omega} \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) p(x') \,dx' = p(x) d_{\nu}(x) + O(\varepsilon^{d + 1});
	\end{equation*}
	and thus by the triangle inequality $\sup_{x \in \Omega}|d_n(x) - d_\nu(x) p(x)| = o_P(\varepsilon^d)$.

	As pointed out previously, if $x' \in \Omega_{\varepsilon}$ then $d_{\nu}(x') = \varepsilon^d$. On the other hand by assumption~\ref{asmp:domain} for any $x' \in \Omega, d_{\nu}(x') \geq \varepsilon^d/\alpha$. Thus 
	\begin{equation*}
	\int_{\Omega} \frac{\eta(\|x' - x\|/\varepsilon)}{d_n(x')} p(x') \,dx' = \int_{\Omega} \frac{\eta(\|x' - x\|/\varepsilon)}{d_{\nu}(x') - o_P(\varepsilon^d)} \,dx' \leq 
	\begin{dcases*}
	\frac{1}{1 - o_P(1)}, & \textrm{if $x \in \Omega_{2\varepsilon}$} \\
	\frac{1}{\alpha} \cdot \frac{1}{1 - o_P(1)}, & \textrm{otherwise.}
	\end{dcases*}
	\end{equation*}
	Plugging back in to~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_2}, we deduce that
	\begin{equation*}
	\|T_{n,\varepsilon}u\|_P^2 \leq \sum_{i = 1}^{n} u_i^2 \biggl(\1\Bigl\{x \in \Omega_{2\varepsilon}\Bigr\} \frac{1}{1 - o_P(1)} + \1\Bigl\{x \in (\partial\Omega)_{2\varepsilon}\Bigr\} \frac{1}{\alpha} \frac{1}{1 - o_P(1)}\biggr).
	\end{equation*}
	This establishes~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_1}, and in turn~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0}.

	Now, suppose $\eta(\cdot)$ can take on negative values. In this case, we have $\eta(\cdot) = \mathrm{sgn}(\eta(\cdot))K^{1/2}(\cdot)$. Let $d_n(x;K^{1/2})$ be the empirical degree with respect to kernel function $K^{1/2}$. Then
	\begin{align*}
	T_{n,\varepsilon}(\wh{\theta} - \varepsilon)(x) & = \frac{1}{d_n(x)n} \sum_{i = 1}^{n} \bigl(\wh{\theta}_i - f^{\ast}(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr) \\
	& = \frac{d_n(x;K^{1/2})}{d_n(x)}\frac{1}{d_n(x;K^{1/2})n} \sum_{i = 1}^{n} \bigl(\wh{\theta}_i - f^{\ast}(X_i)\bigr) \mathrm{sgn}\biggl(\eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr)\biggr) K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
	\end{align*}
	Taking the square and applying Jensen's inequality, we deduce that
	\begin{align*}
	\bigl(T_{n,\varepsilon}(\wh{\theta} - \varepsilon)(x)\bigr)^2 & \leq \biggl(\frac{d_n(x;K^{1/2})}{d_n(x)}\biggr)^2 \cdot \frac{1}{d_n(x;K^{1/2})n}\sum_{i = 1}^{n} \bigl(\wh{\theta}_i - f^{\ast}(X_i)\bigr)^2 K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr) \\
	& \leq \frac{(1 + o_P(1))\|\eta\|_{\infty}}{\alpha} \cdot \frac{1}{d_n(x;K^{1/2})n}\sum_{i = 1}^{n} \bigl(\wh{\theta}_i - f^{\ast}(X_i)\bigr)^2 K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr),
	\end{align*}
	with the latter inequality following by Lemma~\ref{lem:measure_of_ball_estimate}. From here the proof proceeds along similar lines to the case where $\eta$ is non-negative.

	\textit{Proof of~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0.5}.}  
	We will begin by introducing some notation. For a function $u \in L^2(\Omega)$, let
	\begin{equation*}
	L_{\varepsilon,n}u(x) := T_{\varepsilon,n}u(x) - u(x) = \frac{1}{d_n(x)n} \sum_{i = 1}^{n} \bigl(u(x) - u(x_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr)
	\end{equation*}
	and let
	\begin{equation*}
	\wt{L}_{\varepsilon,n}u(x) := \frac{d_n(x)}{d_P(x)}L_{\varepsilon,n}u(x) = \frac{1}{d_P(x)n} \sum_{i = 1}^{n} \bigl(u(x) - u(x_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
	\end{equation*}
	We have already shown that $d_n(x)/d_p(x) = (1 + o_P(1))$, and so we will focus our analysis on upper bounding $\Ebb[\|\wt{L}_{\varepsilon,n}f^{\ast}\|_P^2]$. We begin by providing expressions for the pointwise mean and variance of $\wt{L}_{\varepsilon,n}f^{\ast}(x)$:
	\begin{equation*}
	\Ebb\bigl[\wt{L}_{\varepsilon,n}f^{\ast}(x)\bigr] = T_{P,\varepsilon}f^{\ast}(x) - f^{\ast}(x)
	\end{equation*}
	and
	\begin{align*}
	\mathbb{V}\bigl[\wt{L}_{\varepsilon,n}f^{\ast}(x)\bigr] & = \frac{1}{(d_P(x))^2 n} \mathbb{V}\bigl[\bigl(f^{\ast}(X) - f^{\ast}(x)\bigr) \eta\bigl(\|X - x\|/\varepsilon\bigr)\bigr] \\
	& \leq \frac{1}{p_{\min}^2 \alpha^2 \varepsilon^{2d}} \mathbb{E}\bigl[\bigl(f^{\ast}(X) - f^{\ast}(x)\bigr)^2 K\bigl(\|X - x\|/\varepsilon\bigr)\bigr] \\
	& = \frac{1}{p_{\min}^2 \alpha^2 \varepsilon^{2d}} \int_{\Omega} \bigl(f^{\ast}(x') - f^{\ast}(x)\bigr)^2 K(\|x' - x\|/\varepsilon) p(x) \,dx.
	\end{align*}
	Then, integrating with respect to $P$ over $\Omega$, we obtain
	\begin{align*}
	\Ebb\bigl[\|\wt{L}_{\varepsilon,n}f^{\ast}\|_P^2\bigr]  & \leq \|T_{P,\varepsilon}f^{\star} - f^{\star}\|_{P}^2 + \frac{1}{p_{\min}^2 \alpha^2 \varepsilon^{2d}} E_{\varepsilon}(f^{\ast}) \\
	& \leq \|T_{P,\varepsilon}f^{\star} - f^{\star}\|_{P}^2 + \frac{C\varepsilon^2}{p_{\min}^2 \alpha^2 \varepsilon^{d}} |f|_{H^1(\Omega)}^2
	\end{align*}
	where $E_{\varepsilon}(f^{\ast}) := \int_{\Omega} \int_{\Omega} (f^{\ast}(x') - f^{\ast}(x))^2 K(\|x' - x\|/\varepsilon) p(x') p(x) ,\dx' \,dx$, and we have shown in our AISTATS paper that $E_{\varepsilon}(f^{\ast}) \leq C \varepsilon^{d + 2} \|f\|_{H^1(\Omega)}^2$, justifying the second inequality.
\end{proof}

\begin{lemma}
	\label{lem:measure_of_ball_estimate}
	With probability at least $1 - C \varepsilon^{-d}\exp\{-Cn\varepsilon^d\}$, it holds that
	\begin{equation*}
	\sup_{x \in \Omega}|d_n(x) - d_P(x)| \leq C \sqrt{\frac{\varepsilon^d}{n}},
	\end{equation*}
	and
	\begin{equation*}
	\sup_{x \in \Omega}|d_n(x;K^{1/2}) - d_P(x;K^{1/2})| \leq C \sqrt{\frac{\varepsilon^d}{n}}
	\end{equation*}
\end{lemma}

\begin{proof}
	\textcolor{red}{(TODO)}
\end{proof}

\section{Nystr\"{o}m extension of Laplacian eigenvectors}
The Nystr\"{o}m extension $\varphi_k$ of an eigenvector $v_k$ of the graph Laplacian $L$ is
\begin{equation}
\label{eqn:nystrom}
\varphi_k(x) := \frac{1}{d_n(x) - \lambda_k} \sum_{i = 1}^{n} v_{k,i} K\Bigl(\frac{\|X_i - x\|}{r}\Bigr)
\end{equation}
where $d_n(x) := \sum_{i = 1}^{n} K(\|X_i - x\|/r)$, and we assume $d_n(x) > \lambda_k$ for all $x \in \Omega$ and $k = 1,\ldots,K$. There are certain nice properties of the Nystr\"{o}m extension. It is an interpolant---that is, for any $\theta \in \mathrm{span}\{v_1,\ldots,v_K\}$ the function $f \in \Leb^2(\Omega)$ will agree with $\theta$ at sample design points---and it requires no information on $\Omega$ to be well-defined. However it is also tricky to analyze, because it involves taking $\theta \in \Leb^2(X)$, transforming it into eigenspace, and then applying a \emph{different} smoother to each eigenvector. To build some intuition, we will warm up by analyzing piecewise constant extension, which is much easier to tackle theoretically, if suboptimal for $f^{\ast} \in H^s(\mc{X})$ when $s > 1$.

\subsection{Piecewise constant extension}
Both $\|\Sigma\|$ and $\|f_{\perp}^{\ast}\|_P^2$ are easy to analyze when $\varphi_1,\ldots,\varphi_k$ are taken to be piecewise constant extensions of eigenvectors $v_k$ of the Laplacian matrix $L$.
Let $\{V_1,\ldots,V_n\}$ partition the domain $\Omega$ into $n$ cells, each of probability mass $\Pbb(V_j) = 1/n$, such that $x_j \in V_j$. The piecewise constant extension operator $P^{\star}: L^2(X)\to L^2(\Omega)$ is
\begin{equation}
\label{eqn:piecewise_extension}
\bigl(P^{\star}\theta\bigr)(x) = \sum_{j = 1}^{n} \theta_j \1\{x \in V_j\},
\end{equation}
and we take $\varphi_k = P^{\star}v_k$ for $k = 1,\ldots,K$.

It is clear that $\dotp{{\varphi}_k}{{\varphi}_{\ell}}_{P} = \dotp{v_k}{v_{\ell}}_{n} = \delta_{ij}$, so that $\Sigma = I$ and $\|\Sigma\| = 1$. Moreover, we have that $f_K^{\ast} = P^{\ast}\wt{f}_K$, so that
\begin{equation*}
\|P^{\star}f_{1:n}^{\ast} - {f}_K^{\ast}\|_P = \|f^{\ast} - \wt{f}_K\|_n = \|f^{\ast} - f_K^{\ast}\|_n;
\end{equation*}
the term $\|f^{\ast} - \wt{f}_K\|_n$ is an in-sample bias term that we have already analyzed. It remains to upper bound $\|P^{\star}f^{\ast} - f^{\ast}\|_P$, which we can do using crude but sufficient techniques when $f^{\ast} \in C^1(\Omega)$.

\subsection{Nystr\"{o}m extension}

The analysis of the Nystr\"{o}m extension seems challenging. At present I have suitable bounds on zero out of three of the functionals in question ($\|\Sigma\|$, $\|f_{\perp}^{\ast}\|_n^2$ and $\|f_{\perp}^{\ast}\|_P^2$).

\paragraph{Energy of population-level covariance $(\|\Sigma\|)$.}
The following representation of $\|\Sigma\|$ seems useful:
\begin{equation}
\label{eqn:covariance_energy}
\|\Sigma\| = \sup_{\substack{f \in \mathrm{span}\{\varphi_1,\ldots,\varphi_K\} \\ \|f\|_n = 1}} \|f\|_P.
\end{equation}
The equivalence~\eqref{eqn:covariance_energy} suggests a useful intermediary step in upper bound $\|\Sigma\|$ is an upper bound on $\|\varphi_k\|_P^2$. To do so, it is useful express the Nystrom extension of $v_k$ in terms of a kernel smoother. Recall that $T_{r,n}$ is the kernel smoother,
\begin{equation*}
\bigl(T_{r,n}\theta\bigr)(x) := \frac{1}{d_n(x)} \sum_{i = 1}^{n} \theta_i K\biggl(\frac{\|X_i - x\|}{r}\biggr).
\end{equation*}
Then constructing $\varphi_k$ is equivalent to first passing a kernel smoother over $v_k$, then applying a multiplication operator, i.e
\begin{equation}
\label{eqn:nystrom_extension_kernel_smoother_representation}
\varphi_k(x) = \frac{d_n(x)}{d_n(x) - \lambda_k} \bigl(T_{r,n}v_k\bigr)(x).
\end{equation}  
The operator $T_{r,n}$ is a contraction, in the sense that $\|T_{r,n}\theta\|_P \leq C \|\theta\|_n$, as the following derivations make clear:
\begin{align*}
\|T_{r,n}\theta\|_P^2 & = \int \Biggl(\frac{1}{d_n(x)} \sum_{i = 1}^{n} \theta_i K\biggl(\frac{\|X_i - x\|}{r}\biggr)\Biggr)^2 \,dP(x) \\
& \leq \sum_{i = 1}^{n} \theta_i^2 \int \frac{1}{d_n(x)}  K\biggl(\frac{\|X_i - x\|}{r}\biggr) \,dP(x) \\
& \leq \|\theta\|_n^2 \cdot \frac{d_{\max}(P)}{d_{\min}(P_n)}.
\end{align*}
(One can do better than this, actually, but this suffices for our purposes).

Therefore from~\eqref{eqn:nystrom_extension_kernel_smoother_representation}, and Holder's inequality, we have that
\begin{equation}
\label{eqn:nystrom_extension_l2p_norm}
\|\varphi_k\|_P \leq \sup_{x \in \Omega} \biggl\{\frac{d_n(x)}{d_n(x) - \lambda_k} \biggr\} \cdot  \|T_rv_k\|_P \leq \frac{d_{\min}(P_n)}{d_{\min}(P_n) - \lambda_k/n} \cdot \sqrt{ \frac{d_{\max}(P)}{d_{\min}(P_n)}}.
\end{equation}
For \textcolor{red}{typical} choices of $K$ and $r$, we have that $d_{\min}(P_n) \gg \lambda_K/n$, and therefore the upper bound~\eqref{eqn:nystrom_extension_l2p_norm} is $O(1)$, which suffices for our purposes.

Unfortunately, notwithstanding my prior claim, I don't yet see the right way to leverage this into a bound on $\|\Sigma\|$. For a given $\theta = \sum_{k = 1}^{K} \beta_k v_k$, and the resulting extrapolant $f = \sum_{k = 1}^{K} \beta_k \varphi_k(x)$, we can write $f$ in terms of $T_r\theta$ and a remainder term as follows,
\begin{equation*}
f(x) = \bigl(T_r\theta\bigr)(x) + \sum_{k = 1}^{K} \beta_k \frac{\lambda_k}{d_n(x) - \lambda_k} \bigl(T_rv_k\bigr)(x).
\end{equation*}
However, my best analysis of the remainder term, which I state here, is doomed by accumulation of error:
\begin{align*}
\biggl\|\sum_{k = 1}^{K} \beta_k \frac{\lambda_k}{d_n(x) - \lambda_k} \bigl(T_rv_k\bigr)(x)\biggr\|_P^2 & \overset{(i)}{\leq} K \cdot \sum_{k = 1}^{K} \beta_k^2  \biggl(\frac{\lambda_k/n}{d_{\min}(P_n) - \lambda_k/n}\biggr)^2 \cdot  \|T_rv_k\|_P^2 \\
& \overset{(ii)}{\leq} K \biggl(\frac{\lambda_K/n}{d_{\min}(P_n) - \lambda_K/n}\biggr)^2 \cdot \frac{d_{\max}(P)}{d_{\min}(P_n)} \\
& \asymp \frac{K \lambda_K^2}{n^2d_{\min}(P_n)}.
\end{align*}
To see that this bound is insufficient: recall the optimal choice $K_{\star} \asymp n^{d/(d + 2s)}$, that the eigenvalue $\lambda_K \asymp nr^{d + 2}K^{2/d}$ and that $d_{\min}(P_n) \asymp r^d$. Thus the bound reduces to $n^{\gamma}$ with exponent $\gamma = 3d/(2s + d) - 4/d$, which is greater than $0$ for most values of $d$. Moreover, the looseness in inequality $(ii)$ did not kill us---replacing $\lambda_K$ by $\lambda_2$ on the right hand side of $(ii)$ still results in a bound growing with $n$. On the other hand, the inequality $(i)$ was rather gross---it is tight only when $\varphi_k \propto \varphi_{\ell}$ for all $k,\ell = 1,\ldots,K$, and it seems likely that we can do better.

\paragraph{Distance between $f^{\ast}$ and subspace spanned by $\varphi_1,\ldots,\varphi_K$ $(f_{\perp}^{\ast})$.}

It seems like a useful starting place would be to take $f^{\ast} \propto \psi_k$, where $\psi_k$ is the $k$th eigenvector of the continuum weighted Laplace-Beltrami operator, and show that this is close to $\mathrm{span}\{\varphi_1,\ldots,\varphi_K\}$. I do not know an upper bound on the distance between $\psi_k$ and $\mathrm{span}\{\varphi_1,\ldots,\varphi_K\}$, but I do know such a bound for a different extension of $v_1,\ldots,v_K$.

\paragraph{$(\|f_{\perp}^{\ast}\|_n)$.}

This one I frankly have no thoughts on how to analyze. I don't even understand it all that well, as evidence by the fact that I can't name it.

\clearpage

\section{Prediction error for least-squares estimators: random design, fixed basis.}
In this section, we will assume that $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported over $\Omega \subseteq \Rd$, and that $\phi_1,\ldots,\phi_k$ are deterministic functions which in particular do not depend on $X_1,\ldots,X_n$.

The prediction error of $\wh{f}$ is $\mathbb{E}_{(X,Y)}[(Y - \wh{f}(X))^2]$, where $X \sim P$ and $\mathbb{E}[Y|X] = f^{\ast}(X)$ are independent of $X_1,\ldots,X_n$, and we additionally assume $Y - \mathbb{E}[Y|X]$ is independent of $X$ (i.e. homoskedastic noise). The prediction error can be decomposed into the sum of the $L^2(P)$ error and an irreducible error,
\begin{equation}
\mathbb{E}_{(X,Y)}[(Y - \wh{f}(X))^2] = \|\wh{f} - f^{\ast}\|_P^2 + \mathbb{E}_{(X,Y)}[(Y - f^{\ast}(X))^2],
\end{equation}
justifying our interest in the $\|\cdot\|_P^2$ error.

We will apply Lemma~\ref{lem:l2p_error_fixed_design}, and then use our assumptions on the design points and basis functions to control the right hand side of~\eqref{eqn:l2p_error_fixed_design}. In particular, we have the following. 
\begin{lemma}
	content...
\end{lemma}

\end{document}

### 2021-02-11 notes/manuscripts/laplacian_smoothing/AISTATS/archive/minimax_optimal_laplacian_smoothing.tex
Tampered with directory structure.

--- 2020-09-30 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing2.tex
@@
 \newcommand{\LS}{\mathrm{LS}}
-\newcommand{\SM}{\mathrm{SM}}
 \newcommand{\OS}{\mathrm{OS}}
@@
 \begin{itemize}
-	\item In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Reals^d$, and
+	\item In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Rd$, and
 	\begin{equation}
@@
 	with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise.
-We will assume throughout that $P$ admits a density $p$ with respect to the volume form of $\mc{X}$ which is bounded away from $0$ and $\infty$, i.e.
-	\begin{equation*}
-	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty,~~\textrm{for all $x \in \Xset$.}
-	\end{equation*}
 Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present.
@@
 
-\textbf{(1.a) Motivation.} Laplacian smoothing has many advantages compared to other nonparametric regression methods. For instance:
+\textbf{(2.a)} Laplacian smoothing has many advantages compared to other nonparametric regression methods. For instance:
 
@@
 
-\textbf{(1.b) Related work.} For this reason a considerable body of work has emerged analyzing the convergence of the graph Laplacian $\Lap_{{n,r}}$ to the continuum Laplace operator $\Delta_P: C^2(\Xset) \to \Leb^2(\Xset)$,
-\begin{equation}
-\label{eqn:laplace_operator}
-\Delta_Pg := -\frac{1}{p} \dive(p^2 \nabla g)
-\end{equation}
-as $n \to \infty$ and $r(n) \to 0$. This is meant in various senses:
+\textbf{(2.b)} For this reason a considerable body of work has emerged analyzing the consistency of the graph Laplacian $\Lap_{{n,r}}$ as $n \to \infty$ and $r(n) \to 0$.
+This is meant in various senses:
 \begin{itemize}
-	\item Pointwise convergence, meaning $\Lap_{n,r}f \to \Delta_Pf$. 
-	\item Spectral convergence, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$.
-	\item Convergence of norms, meaning $f^T \Lap_{n,r} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$. 
+	\item Pointwise consistency, meaning $\Lap_{n,r}f \to \Delta_Pf$ for an appropriate limiting operator $\Delta_P$.
+	\item Spectral consistency, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$.
+	\item Consistency of norms, meaning $f^T \Lap_{n,f} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$. 
 \end{itemize}
@@
 
-\textbf{(1.c) Our contributions.} Our main contributions lie in this latter direction, and they can all be summarized as follows: Laplacian smoothing methods are minimax optimal over Sobolev spaces.
+\textbf{(2.c)} Our main contributions lie in this latter direction, and they can all be summarized as follows: Laplacian smoothing methods are minimax optimal over Sobolev spaces.
 In more detail, we will show that when $f_0$ belongs to the Sobolev ball $H^1(\Xset;M)$:
@@
 
-\textbf{(1.d) Organization.}
+\paragraph{Organization.}
 
-\textcolor{red}{(TODO)}
-
-\textbf{(1.e) Notation.}
-
-\textcolor{red}{(TODO)}
+List out what is to come.
 
@@
 
-\textbf{(2)} Before we get to our main results, we briefly review minimax optimal estimation and testing rates over Sobolev classes.
+Before we get to our main results, we briefly review minimax optimal estimation and testing rates over Sobolev classes.
 We also single out a notable method which achieves these rates: \emph{smoothing splines}.
@@
 
-\paragraph{(2.a) Sobolev spaces.} We start by briefly introducing Sobolev spaces, emphasizing only the material relevant to our statistical context. 
-\begin{itemize}
-	\item Roughly speaking, Sobolev spaces contain functions $f \in \Leb^p(\Xset)$ with derivatives which themselves belong to $\Leb^p(\Xset)$.
+\subsection{Sobolev spaces}
+\label{subsec:sobolev_spaces}
+
+We start by briefly introducing Sobolev spaces, emphasizing only the material relevant to our statistical context.
+Roughly speaking, Sobolev spaces contain functions $f \in \Leb^p(\Xset)$ with derivatives which themselves belong to $\Leb^p(\Xset)$.
 More formally, for given integers $s$ and $p > 0$, the Sobolev space $W^{s,p}(\Xset)$ consists of all functions $f \in \Leb^p(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ exists and belongs to $\Leb^p(\Xset)$.
@@
 and for a given $M > 0$, the $(s,p)$-Sobolev ball is $W^{s,p}(\Xset, M) = \set{f: \norm{f}_{W^{s,p}(\Xset)} \leq M}$.
-In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space, and we adopt the usual convention of writing $H^s(\Xset) = W^{s,2}(\Xset)$ and $H^s(\Xset,M) = W^{s,2}(\Xset,M)$.
-Linear estimators such as $\wh{f}_{\LS}$\footnote{To be explicit, estimators of the form $\wh{f} = HY$ for some matrix $H \in \Reals^{n \times n}$.} can be very accurate over the Hilbert-Sobolev spaces $H^s(\Xset)$, and we will confine our attention to this case hereafter.
-	\item \textcolor{red}{(TODO)}: Introduce the Sobolev embedding theorem, for at least three reasons: (1) to facilitate later discussion of why smoothing splines are ill-posed when $2s < d$; (2) to make explicit why the statistical problem isn't ill-posed when $2s < d$ in the random design setting; (3) to make clear why the testing problem is hopeless when $4s < d$ without imposing additional conditions.
-\end{itemize}
+In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space, and we adopt the usual convention of writing $H^s(\Xset) = W^{s,2}(\Xset)$ and $H^s(\Xset,M) = W^{s,2}(\Xset,M)$; we will confine our attention to this case hereafter.
 
-\paragraph{(2.b) Connecting Laplacian smoothing to smoothing splines.} 
-To better motivate Laplacian smoothing as a tool well-suited for nonparametric regression, we review their connection to one of the most classical methods of nonparametric regression: smoothing splines.
-\begin{itemize}
-	\item Smoothing splines are particularly well-suited for nonparametric regression over Sobolev spaces.
-This is because smoothing splines are penalized least squares estimators where the penalty is itself the Sobolev semi-norm.
-Mathematically, the smoothing spline estimator $\wh{f}_{\SM}$ is defined as the solution to the following optimization problem,
-	\begin{equation}
-	\label{eqn:smoothing_spline}
-	\wt{f} = \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \cdot \seminorm{f}_{H^{1}(\Xset)}^2\biggr\}.
-	\end{equation}
-	In~\eqref{eqn:smoothing_spline}, the minimum can either be taken over all $f \in H^1(\Xset)$ or over all $f \in C^1(\Xset)$; we shall refer to the former as smoothing splines, and the latter as \emph{Holder} smoothing splines.
+\textbf{(4)} In fact, an alternative way to motivate Laplacian smoothing nonparametric regression methods---as opposed to the advantages mentioned in \textbf{(2.a)}---comes from viewing them as discrete-time versions of smoothing splines.
+This is made more rigorous by the consistency results of \textbf{(2.b)}.
+Viewed in this light, an additional advantage of Laplacian smoothing emerges: it automatically adapts to the geometry of $\Xset$ and $p$.
 
-	\textcolor{red}{(TODO)}: Check the classical spline references, and/or Ryan, for data on (1) what the canonical domain of minimization is for smoothing splines, and (2) whether the representer theorem depends holds for either domain of minimization.
-
-	\item By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can tie the penalty used by the smoothing spline estimator $\wh{f}_{\SM}$ to that used by the Laplacian smoothing estimator $\wh{f}_{\LS}$.
-Suppose for the moment that $P$ is the uniform distribution over $\Xset$.
-Using the definition~\eqref{eqn:laplace_operator}, integrating by parts, and invoking the aforementioned convergence properties of $\Lap_{n,r}$, we have that for any $f \in C^{2}(\Xset)$ which satisfies appropriate boundary conditions,
-	\begin{equation}
-	\label{eqn:seminorm_convergence}
-	\abs{f}_{H^1(\Xset)} = \int_{\Xset} \bigl(\Delta_P f\bigr)(x) \cdot f(x) \,dx = \lim_{n \to \infty} \frac{1}{n^{2}r^{(d + 2)}}f^T \Lap_{n,r} f.
-	\end{equation}
-	\textcolor{red}{(TODO)}: Formalize the boundary conditions---I believe either Neumann or Dirichlet boundary conditions should suffice?
-	\item \eqref{eqn:seminorm_convergence} establishes asymptotic equality (up to a scaling factor) between the penalties in~\eqref{eqn:smoothing_spline} and~\eqref{eqn:laplacian_smoothing}, when the distribution $P$ is uniform.
-But it also sheds light on additional feature of Laplacian smoothing, relevant when $P$ is non-uniform: the penalty $f^T \Lap_{n,r} f$ automatically adapts to the design density $p$, in a way that the penalty $|f|_{H^1(\Xset)}$ does not.
-This automatic adaptivity to $P$ turns out to be a formal advantage of Laplacian smoothing, which we will comment on further in Section~\ref{sec:minimax_optimal_laplacian_smoothing}.
-\end{itemize}
-
-\paragraph{(2.c) Minimax-optimality of smoothing splines.} 
-One reason all of this is intriguing is that, as mentioned previously, the smoothing spline estimator $\wh{f}_{\SM}$---and the resultant test statistic $T_{\SM} := \|\wt{f}_{\SM}\|_n^2$--- have strong statistical properties.
-To be precise, they achieve the well-known minimax rates over Sobolev spaces.
-\begin{itemize}
-	\item  The minimax estimation rate over Sobolev balls, with error measured in $\Leb^2(\Xset)$ norm, is 
-	\begin{equation}
-	\label{eqn:sobolev_space_estimation_minimax_rate}
-	\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{d/(2 + d)}n^{-2/(2 + d)}~~\textrm{for all $d \geq 1$.}
-	\end{equation}
-
-	\textcolor{red}{(TODO)}: (1) You need a reference for this statement, and you already know one will be tricky to find because of ill-definedness of point evaluation.
-(2) This statement only holds under some condition on the density $p$.
-
-	\item On the other hand, the minimax critical radius over Sobolev balls---that is, the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has non-trivial power over all $H_{\epsilon}^1(\Xset,M) := H^1(\Xset,M) \cap \{f: \norm{f}_{\Leb^2(\Xset)} \geq \epsilon\}$---is much smaller,
-	\begin{equation}
-	\label{eqn:sobolev_space_testing_critical_radius}
-	\epsilon_n^{\star}\Bigl(H^1(\Xset,M)\Bigr):= \inf\Biggl\{\epsilon: \inf_{\phi} \biggl[\Ebb_0[\phi] +  \sup_{f_0 \in H_{\epsilon}^1(\Xset,M)} \Ebb_{f_0}[1 - \phi]\biggr] \leq 2\alpha \Biggr\} \asymp n^{-2/(4 + d)}~~\textrm{for $1 \leq d < 4$,}
-	\end{equation}
-
-	\textcolor{red}{(TODO)}: (1) You need a reference for this statement (Ingster), and (2) you need to explain what happens when $d \geq 4$.
-
-	\item The estimator $\wh{f}_{\SM}$ and the level-$\alpha$ test $\phi_{\SM} := \1\{T_{\SM} \geq {t}_{\alpha}\}$ are known to be minimax rate-optimal---meaning they achieve the rates in~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius}, respectively---when $d = 1$.
-When $d > 1$, as a consequence of the Sobolev Embedding Theorem\footnote{Or to be more precise, as a consequence of the \emph{tightness} of the Sobolev Embedding Theorem.} the variational problem~\eqref{eqn:smoothing_spline} is not well-posed: the criterion is minimized by interpolating the responses $\mathbf{Y}$, and the resulting estimator $\wt{f}$ will be inconsistent.
-By replacing the domain of minimization $H^1(\Xset)$ by $C^1(\Xset)$, we can force the solution to~\eqref{eqn:smoothing_spline} to be less wild, and the resulting Holder smoothing spline is not merely consistent but in fact minimax optimal when $d = 2$.\footnote{At least in the estimation problem.
-We could not find either a positive or negative result for the testing problem.} However when $d > 2$, the Holder space $C^1(\Xset)$ is itself too large, and even the Holder smoothing spline has risk provably larger than optimal.
-Previewing things a bit, we will see in Section~\ref{sec:minimax_optimal_laplacian_smoothing} that a similar phenomenon occurs for the Laplacian smoother $\wh{f}$; interestingly, the breakdown point will come when $d > 4$, rather than $d > 1$ or $d > 2$.
-
-	\textcolor{red}{(TODO)}: Include references.
-
-	\item Finally, it is well known that if $\Xset$ satisfies the manifold hypothesis, meaning informally that it has intrinsic dimension $m < d$ (see~\ref{asmp:domain_manifold} in Section~\ref{sec:graph_sobolev_classes} for a formal statement), both~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
-Encouragingly, the smoothing spline estimator $\wt{f}$ and test $\wt{\phi}$ achieve these faster rates, with one caveat: in order for smoothing splines to properly leverage the manifold hypothesis, the domain $\Xset$ must be known.
-\end{itemize}
-
-\paragraph{(2.d) Consequences?} Intuitively, the connections between Laplacian smoothing and smoothing splines give us reason to hope that the strong statistical properties of the latter might be shared by the former.
-Unfortunately, it is not so easy to mathematically tie $\wh{f}$ to $\wt{f}$: for example, note that the convergence in~\eqref{eqn:seminorm_convergence} assumes $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:smoothing_spline} is either $H^1(\Xset)$ or $C^1(\Xset)$.
-Instead, we will take a different approach and analyze $\wh{f}$ by leveraging properties of the graph Sobolev class $H^1(G_{n,r})$, which we now introduce.
+\textbf{(5)} Smoothing splines are known to have strong minimax optimal properties. Reasoning by analogy, we might hope that Laplacian smoothing would share these properties.
 
 \section{Graph Sobolev classes}
-\label{sec:graph_sobolev_classes}
 
-\textbf{(3)}  For $M > 0$, define the (first-order) graph Sobolev ball to be
-\begin{equation}
-\label{eqn:graph_sobolev_ball}
-H^1(G_{n,r},M) = \Bigl\{f \in \Reals^n: f^T \Lap_{n,r} f \leq M^2\Bigr\}
-\end{equation}
-where we recall that the \emph{graph Sobolev seminorm} $f^T \Lap_{{n,r}} f$ is precisely the penalty term in~\eqref{eqn:laplacian_smoothing}.
+\textbf{(6)} The minimax optimality of smoothing splines rests on two crucial facts: one, the a priori assumption that $|f|_{H^1(\Xset)}$ is bounded; two, an upper bound on the metric entropy of $H^1(\Xset)$, characterized by the decay of its eigenvalues.
+\textcolor{red}{(TODO)}:(I know this isn't well-stated, and I need to clean it up.)  We now show that qualitatively similar statements hold with respect to
 
-To understand why graph Sobolev balls play a critical role in the analysis of Laplacian smoothing, let us return for a moment to the continuous-time setting.
-Upper bounds on the error of smoothing splines rest on two crucial facts, both of which involve the continuum Laplace operator $\Delta_P$.
-The first is the \textit{a priori} assumption that the Sobolev semi-norm $|f|_{H^1(\Xset)} = \langle f, \Delta_P f \rangle_{\Leb^2(P)} \leq M^2$.\footnote{The equality holding when $P$ is uniform and  only for functions $f$ which satisfy boundary conditions.} The second is an upper bound on the metric entropy of $H^1(\Xset)$, characterized by the eigenvalue decay $\lambda_k(\Delta_P) \asymp k^{2/d}$ (i.e.
-Weyl's Law).
-In this section we will show that both these statements have analogues in the graph setting:
-\begin{itemize}
-	\item for functions $f$ belonging to $H^1(\Xset,M)$, the evaluations $f \in (f(X_1),\ldots,f(X_n)) \in \Reals^n$ are upper bounded,
-	\begin{equation}
-	\label{eqn:graph_sobolev_seminorm_vague}
-	f^T \Lap_{n,r} f \leq C M^2(n^{2}r^{(d + 2)}),
-	\end{equation}
-	\item for a range $1 \leq k \leq k_{\max}(n)$, the eigenvalues $\lambda_k(G_{n,r})$ satisfy
-	\begin{equation}
-	\label{eqn:neighborhood_graph_eigenvalue_vague}
-	ck^{2/d} \leq \frac{1}{nr^{d + 2}} \lambda_k(G_{n,r})  \leq Ck^{2/d}.
-	\end{equation}
-\end{itemize}
-Both statements will hold with high probability, and when we assume $\Xset$ is a manfiold of dimension $m$, they will hold with $d$ replaced by $m$.
-These results elucidate why Laplacian smoothing ``works'' as well as its continuous-time counterpart ---in the sense of being statistically rate-optimal over (continuum) Sobolev classes.
+\quad \textbf{(6.1)} the graph Sobolev semi-norm $f_0^T \Lap_{n,r} f_0$ and,
 
-\paragraph{(3.a) Graph Sobolev semi-norm.}
-We begin by upper bounding $f_0^T \Lap_{n,r} f_0$ under the assumption $f_0 \in H^1(\Xset,M)$. We will assume the kernel $K$ satisfies condition~\ref{asmp:kernel}.
-\begin{enumerate}[label=(K\arabic*)]
-	\item
-	\label{asmp:kernel}
-	$K:[0,\infty) \to [0,\infty)$ is supported on $[0,1]$, and its restriction to $[0,1]$ is Lipschitz. Additionally, it is normalized so that
-	\begin{equation*}
-	\int_{\Reals^d} K\bigl(\norm{z}\bigr) \,dz = 1.
-	\end{equation*}
-\end{enumerate}
-This is a relatively mild assumption: the choice of kernel is under the user's control, and moreover~\ref{asmp:kernel} covers many (though certainly not all) common kernel choices.
-In Lemma~\ref{lem:graph_sobolev_seminorm}, we prove that if the density $p$ is merely upper bounded then~\eqref{eqn:graph_sobolev_seminorm_vague} holds with respect to the graph Sobolev semi-norm.
-\begin{lemma}
-	\label{lem:graph_sobolev_seminorm}
-	Assume~\ref{asmp:kernel} and additionally that $p(x) \leq p_{\max}$ for all $x \in \Xset$.
-Then, for any $f \in H^1(\Xset)$, there exists a constant $C_1 > 0$ which depends only on $\Xset$, $d$, $p_{\max}$ and $K$ such that
-	\begin{equation}
-	\label{eqn:graph_sobolev_seminorm}
-	f^T \Lap_{n,r} f \leq \frac{C_1}{\delta} n^2 r^{d + 2} |f|_{H^1(X)}^2
-	\end{equation}
-	with probability at least $1 - \delta$.
-\end{lemma}
-A few remarks:
-\begin{itemize}
-	\item 
-	The inequality~\eqref{eqn:graph_sobolev_seminorm} resembles~\eqref{eqn:seminorm_convergence}.
-However, as we pointed out previously to derive~\eqref{eqn:seminorm_convergence} one needs $f \in C^{2}(\Xset)$ as opposed to $f \in H^1(\Xset)$, and imposing this extra degree of regularity is unsuitable for our purposes.
-We use a more careful analysis to prove Lemma~\eqref{lem:graph_sobolev_seminorm}, but note well that~\eqref{eqn:graph_sobolev_seminorm} is only an upper bound rather than a statement about consistency.
-This is the price we pay for insisting that $f$ have only one derivative; fortunately an upper bound will be sufficient for our purposes in Section~\ref{sec:minimax_optimal_laplacian_smoothing}.
-	\item We note also that~\eqref{eqn:graph_sobolev_seminorm}---which holds with probability $1 - \delta$---depends proportionally on $1/\delta$, as opposed to $\log(1/\delta)$.
-The reason: the Sobolev assumption $f \in H^1(\Xset)$ only grants us control over the second moment of $f$ and its first derivative, and so we are afforded only the relatively weak concentration implied by Markov's inequality.
-If we assume $f$ and its first derivative are more uniformly bounded--for instance if $f \in C^1(\Xset)$---we can use Hoeffding's inequality to get sharper upper bounds.
-\end{itemize}
-
-\textbf{(3.b) Graph Laplacian eigenvalues.} The graph eigenvalues $\lambda_k(\Lap_{n,r}) \asymp n r^{d + 2}k^{2/d}$---when $\Xset$ is full-dimensional---or $\lambda_k(\Delta_P) \asymp nr^{m + 2}k^{2/m}$---when $\Xset$ has intrinsic dimension $1 < m < d$.
+\quad \textbf{(6.2)} the graph eigenvalues $\lambda_k(\Lap_{n,r})$.
 
 \section{Minimax optimality of Laplacian smoothing}
-\label{sec:minimax_optimal_laplacian_smoothing}
 
-\textbf{(4)} We now leverage the results of Section~\ref{sec:graph_sobolev_classes} to formalize the main conclusions of this work: that Laplacian smoothing methods are minimax rate-optimal.
+\textbf{(7)} Using \textbf{(6.1)} and \textbf{(6.2)}, we establish upper bounds on the estimation and testing error of Laplacian smoothing, which confirm that Laplacian smoothing is asymptotically minimax rate-optimal over the Sobolev class $H^1(\Xset)$.
 
-\textbf{(4.a) Estimation error of Laplacian smoothing.} Using \textbf{(3.a)} and \textbf{(3.b)}, we establish upper bounds on the estimation error of Laplacian smoothing.
+\textbf{(8)} While we motivated graph Laplacians by suggesting they might replicate the statistical properties of smoothing splines, we have in fact managed to show something stronger. In particular:
 
-\textbf{(4.b) Testing error of Laplacian smoothing.} Using \textbf{(3.a)} and \textbf{(3.b)}, we establish upper bounds on the testing error of Laplacian smoothing.
+\quad \textbf{(8.1)} When $d \geq 2$, there is no Sobolev embedding of $H^1(\Xset) \subseteq C^s(\Xset)$ for any $s > 0$. Therefore the smoothing spline estimator is ill-posed.
 
-\quad \textbf{(4.b.i) Low-smoothness regime.} Unlike in the estimation setting, the nonparametric goodness-of-fit testing problem itself exhibits an elbow when $d = 4$ and $f_0 \in \Leb^4(\Xset)$.
-The naive statistic $\frac{1}{n}\|Y\|_2^2$ is now optimal.
-By tuning Laplacian smoothing to the interpolation limit---i.e.
-setting $\rho = 0$---we recover this naive statistic, and therefore Laplacian smoothing is optimal in this sense (although really, no smoothing is being done.)
-
-\textbf{(4.c) Comparing with smoothing splines.} While we motivated graph Laplacians by suggesting they might replicate the statistical properties of smoothing splines, we have in fact managed to show something stronger.
-In particular:
-
-\quad \textbf{(4.c.i)} When $d \geq 2$, there is no Sobolev embedding of $H^1(\Xset) \subseteq C^s(\Xset)$ for any $s > 0$. Therefore the smoothing spline estimator is ill-posed.
-
-\quad \textbf{(4.c.ii)} Additionally, the metric entropy of $C^1(\Xset)$ is too large to obtain optimal rates for \textcolor{red}{Holder-smoothing splines} when $d \geq 2$.
+\quad \textbf{(8.2)} Additionally, the metric entropy of $C^1(\Xset)$ is too large to obtain optimal rates for \textcolor{red}{Holder-smoothing splines} when $d \geq 2$.
 Thus, we see a gap emerge between the optimality properties of the discrete-time (Laplacian smoothing) and continuous-time (smoothing splines) estimators, in favor of the latter.
 
-\quad \textbf{(4.c.iii)} On the other hand, when $d \geq 4$, we no longer get the known optimal rates in the estimation problem.
+\quad \textbf{(8.3)} On the other hand, when $d \geq 4$, we no longer get the known optimal rates in the estimation problem.
 When $d = 4$, our rates are suboptimal by a factor of $\log n$.
@@
 We believe this is related to classical results regarding the entropy of low-smoothness Holder classes, i.e.
-the results used to justify \textbf{(4.c.ii)}.
-The interesting point is that the parameter spaces $H^1(\Xset)$, $C^1(\Xset)$ and $H^1(G_{n,r})$ become ``too large'' for different values of $d$.
+the results used to justify \textbf{(8.2)}.
+The interesting point is that the parameter spaces $C^1(\Xset)$ and $H^1(G_{n,r})$ become ``too large'' for different values of $d$.
 
-\quad \textbf{(4.c.iv)} \textcolor{red}{(TODO)}: Translate these results into error in $\Leb^2(P)$ error, if/when possible.
+\quad \textbf{(8.4)} \textcolor{red}{(TODO)}: Translate these results into error in $\Leb^2(P)$ error, if possible.
 
-\textbf{(4.d) Manifold adaptivity.} In Section~\ref{sec:graph_sobolev_classes}, we established that when $\Xset$ is an $m$-dimensional submanifold of $\Reals^d$, Laplacian smoothing approaches automatically ``feel'' the intrinsic dimension $m$ of $\Xset$.
-We now show that this translates into improved upper bounds on estimating and testing error.
+\subsection{Manifold adaptivity}
 
-\quad \textbf{(4.d.i)} This confirms the statement in \textbf{(4)} that Laplacian smoothing automatically adapts to the geometry of $\Xset$ and $p$.
+\quad \textbf{(9)} We now show that when $\Xset$ is an $m$-dimensional submanifold of $\Reals^d$, Laplacian smoothing approaches automatically ``feel'' the intrinsic dimension $m$ of $\Xset$, which translate into improved upper bounds on estimating and testing error.
+This confirms the statement in \textbf{(4)} that Laplacian smoothing automatically adapts to the geometry of $\Xset$ and $p$.
 
-\quad \textbf{(4.d.ii)} By contrast, smoothing splines must be specially tailored to work in this setting.
+\subsection{Higher-order smoothness classes}
 
-\textbf{(4.e) Higher-order smoothness classes.} Laplacian smoothing can also be adapted to take advantage of additional assumed regularity on $f_0$, i.e.
+\quad \textbf{(10)} Analogously to smoothing splines, Laplacian smoothing can also be adapted to take advantage of additional assumed regularity on $f_0$, i.e.
 $f_0 \in H^s(\mc{\Xset})$ for $s > 1$.
-In the very special case of $s = 2$ and $d = 2$, we can show that this adapted estimator achieves the sharper minimax rates over these classes, but the general story--for all combinations of $s$ and $d$---remains beyond our reach.
+In very limited circumstances, we can show that this adapted estimator achieves the sharper minimax rates over these classes.
+However, the general story--for all combinations of $s$ and $d$---remains beyond our reach.
 
@@
 
-\textbf{(5)} Empirically, we demonstrate that the risk for Laplacian smoothing is comparable to that of smoothing splines---in the full dimensional case with uniformly random design.
-When the design is sampled from a density concentrated near a manifold, Laplacian smoothing outperforms (out of the box) smoothing splines.
+\quad \textbf{(11)} Empirically, we demonstrate that the risk for Laplacian smoothing is comparable to that of smoothing splines, in the full dimensional case.
+Although one cannot carry out a comparison in the case of an unknown manifold --- since the smoothing spline estimator must be changed in a non-obvious manner --- we do show that when the density $p$ is concentrated close to a manifold and $f_0$ is smooth where $p$ is peaked, we have improved performance for $\wh{f}_{\LS}$ compared to $\wh{f}_{\mathrm{SM}}$.
 
@@
 
-\textcolor{red}{(TODO)}
-
 \clearpage
-
-\appendix
-
-\clearpage
-
-Many of our results follow the same general two-part proof strategy.
-First, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to functionals of the graph $G$.
-We then analyze the behavior of these functionals with respect to the particular neighborhood graph $G_{n,r}$ and give high probability (upper or lower) bounds on these functionals.
-It is in this second step where we invoke our various assumptions on the distribution $P$ and regression function $f_0$.
-
-\section{Graph-dependent error bounds}
-\label{sec:fixed_graph_error_bounds}
-
-\section{Neighborhood graph Sobolev semi-norm}
-\label{sec:graph_sobolev_seminorm}
-
-\section{Bounds on neighborhood graph eigenvalues}
-\label{sec:graph_eigenvalues}
-
-\section{Bounds on the empirical norm}
-\label{sec:empirical_norm}
-
-\section{Proof of theorems}
 

### 2021-02-11 notes/manuscripts/laplacian_smoothing/AISTATS/archive/minimax_optimal_laplacian_smoothing2.tex
Tampered with directory structure.

--- 2020-09-30 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing2.tex
@@
 \begin{document}
-\title{Minimax optimal Laplacian smoothing}
-\author{Alden Green}
-\date{\today}
-\maketitle
 
-\textbf{Disclaimer:} This largely follows the same setup as in the draft I have written which combines Laplacian smoothing and Laplacian eigenmaps results.
-That draft --- document name \textit{graph\_regression2} --- should for the moment be considered obsolete.
+\twocolumn[
+
+	\aistatstitle{Minimax optimal Laplacian smoothing}
+
+	\aistatsauthor{Alden Green, Sivaraman Balakrishnan, and Ryan J. Tibshirani}
+
+	\aistatsaddress{Carnegie Mellon University}]
+
+\begin{abstract}
+	\textcolor{red}{(TODO)}:
+\end{abstract}
 
@@
 \end{itemize}
+\textcolor{red}{(TODO)}: Perhaps additional motivation is needed?
 
-\textbf{(1.b) Related work.} For this reason a considerable body of work has emerged analyzing the convergence of the graph Laplacian $\Lap_{{n,r}}$ to the continuum Laplace operator $\Delta_P: C^2(\Xset) \to \Leb^2(\Xset)$,
+\paragraph{Related work.} For this reason a considerable body of work has emerged analyzing the properties of Laplacian smoothing, as well as of graph-based methods more generally.
+Roughly speaking this work can divided into two categories.
+
+Works in the first category view the responses $Y = \theta + \varepsilon$ as noisy samples of a regression function $\theta \in \Reals^n$ defined on the vertices of a \emph{fixed graph} $G$.
+and bound the statistical error as a function of $\theta$ and $G$.
+(In the estimation literature, see for instance \citep{wang2016, sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}; in the testing literature, see \citep{sharpnack2013,sharpnack2013b,sharpnack2015}.) Note that in our setting, the graph $G_{n,r}$ and the evaluations $(f_0(X_1),\ldots,f_0(X_n))$ are random rather than fixed, and our assumptions are placed only on the underlying objects $P$ and $f_0$ from which these random object arise.
+As such, both our analysis and conclusions are naturally quite different.
+
+Works in the second category prove the consistency of neighborhood graph Laplacian smoothing and other neighborhood graph-based methods in the large-sample limit.
+They do so by establishing the convergence of the graph Laplacian $\Lap_{{n,r}}$ to the continuum Laplace operator $\Delta_P: C^2(\Xset) \to \Leb^2(\Xset)$,
 \begin{equation}
@@
 \end{equation}
-as $n \to \infty$ and $r(n) \to 0$. This is meant in various senses:
+as $n \to \infty$ and $r \to 0$.
+This is meant in (at least) three senses: (1) \emph{pointwise convergence}, meaning that $\Lap_{n,r}f \to \Delta_Pf$ in an appropriate norm \citep{belkin03,belkin05,lafon04,hein05,singer06,gine06}; (2) spectral convergence, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$, where the latter are the Dirichlet eigenvalues of $\Delta_P$ \citep{belkin07,burago2014,trillos2018,trillos2019,calder2019}.\footnote{\label{footnote:dirichlet_eigenvalues} Formally, Dirichlet eigenvalues are solutions to the problem
+	$$
+	\Delta_Pg = \lambda g,\\
+	\restr{g}{\partial \Xset} = 0
+	$$}, and (3) convergence of norms, meaning $f^T \Lap_{n,r} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$ \citep{bousquet03,hein06,zhou11}.
+While these results imply \emph{consistency} of many different graph Laplacian based procedures, the statistical \emph{optimality} of Laplacian smoothing over neighborhood graphs is still not well understood.
+
+\paragraph{Our contributions.} Our three main contributions lie in this latter direction, and they can all be summarized as follows: Laplacian smoothing methods are minimax optimal over first order Sobolev spaces.
+In more detail, we will show that when $f_0$ belongs to the first order Sobolev ball $H^1(\Xset;M)$:
 \begin{itemize}
-	\item Pointwise convergence, meaning $\Lap_{n,r}f \to \Delta_Pf$. 
-	\item Spectral convergence, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$.
-	\item Convergence of norms, meaning $f^T \Lap_{n,r} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$. 
-\end{itemize}
-However, the statistical optimality of Laplacian smoothing is still not well understood.
-
-\textbf{(1.c) Our contributions.} Our main contributions lie in this latter direction, and they can all be summarized as follows: Laplacian smoothing methods are minimax optimal over Sobolev spaces.
-In more detail, we will show that when $f_0$ belongs to the Sobolev ball $H^1(\Xset;M)$:
-\begin{itemize}
-	\item \textbf{Minimax optimal estimation.} With high probability, $\frac{1}{n}\norm{\wt{f}_{\LS} - f_0}_{2}^2 \lesssim n^{-2/(2 + d)}$.
+	\item \textbf{Minimax optimal estimation.} With high probability, $\frac{1}{n}\bigl\|\wt{f}_{\LS} - f_0\bigr\|_{2}^2 \lesssim M^{d/(2+d)}n^{-2/(2 + d)}$.
 	\item \textbf{Minimax optimal testing.}
-	A level-$\alpha$ test constructed using $T_{\LS}$ has non-trivial power whenever $\norm{f_0}_{\Leb^2(\Xset)}^2 \gtrsim n^{-4/(4 + d)}$. 
+	A level-$\alpha$ test constructed using $T_{\LS}$ has non-trivial power whenever $\norm{f_0}_{\Leb^2(\Xset)}^2 \gtrsim M^{2d/(4 + d)}n^{-4/(4 + d)}$. 
 	\item \textbf{Manifold adaptivity.}
@@
 \end{itemize}
-Finally, we will also establish that the Laplacian smoothing estimator can be altered to take advantage of higher-order smoothness assumptions on $f_0$.
-This last conclusion will hold only under very limited circumstances.
 
-\textbf{(1.d) Organization.}
+\paragraph{Organization.} We now describe the structure of the rest of this paper, which builds to the aforementioned minimax results.
+\begin{itemize}
+	\item In Section~\ref{sec:minimax_optimal_regression_sobolev_spaces}, for completeness we review some standard results regarding minimax nonparametric regression over Sobolev spaces.
+	\item In Section~\ref{sec:graph_sobolev_classes}, we introduce the graph Sobolev class $H^1(G_{n,r})$, which we view as a graph-based analog to the continuum Sobolev class $H^1(\Xset)$.
+We show that key properties of $H^1(\Xset)$ used to derive minimax rates are shared by $H^1(G_{n,r})$.
+	\item In Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we leverage the results of Section~\ref{sec:graph_sobolev_classes} to derive tight upper bounds on the error of Laplacian smoothing methods.
+	\item In Section~\ref{sec:simulations} we provide some experiments supporting our theory, before concluding in Section~\ref{sec:discussion}.
+\end{itemize}
 
-\textcolor{red}{(TODO)}
+\textbf{Notation.}
 
-\textbf{(1.e) Notation.}
+We use $A,A_1, \ldots$ and $C,C_1, \ldots$ to refer to constants which do not depend on $n$, $r$, or $f_0$; likewise with $a, a_1,\ldots$ and $c, c_1,\ldots$.
 
-\textcolor{red}{(TODO)}
+\textcolor{red}{(TODO)}: Complete this.
 
 \section{Minimax-optimal regression over Sobolev spaces}
+\label{sec:minimax_optimal_regression_sobolev_spaces}
 
-\textbf{(2)} Before we get to our main results, we briefly review minimax optimal estimation and testing rates over Sobolev classes.
+Before we get to our main results, we review minimax optimal estimation and testing rates over Sobolev classes.
 We also single out a notable method which achieves these rates: \emph{smoothing splines}.
@@
 
-\paragraph{(2.a) Sobolev spaces.} We start by briefly introducing Sobolev spaces, emphasizing only the material relevant to our statistical context. 
+\paragraph{Sobolev spaces, and the Sobolev embedding theorem.} We start by briefly introducing Sobolev spaces, emphasizing only the material relevant to our statistical context.
+
+Roughly speaking, (Hilbert-)Sobolev spaces contain functions $f \in \Leb^2(\Xset)$ with derivatives which themselves belong to $\Leb^2(\Xset)$.
+More formally, for a given integer $s$, the Sobolev space $H^{s}(\Xset)$ consists of all functions $f \in \Leb^2(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ exists and belongs to $\Leb^2(\Xset)$.
+For such functions, the Sobolev seminorm $\seminorm{f}_{H^{s}(\Xset)}$ and norm $\norm{f}_{H^{s}(\Xset)}$ are given by
+\begin{equation*}
+\seminorm{f}_{H^s(\Xset)}^2 = \sum_{\abs{\alpha} = s}\int_{\mathcal{X}} \Bigl|\bigl(D^{\alpha}f\bigr)(x)\Bigr|^2 \,dx, ~~ \norm{f}_{H^{s}(\Xset)}^2 = \sum_{k = 0}^{s} \seminorm{f}_{H^k(\Xset)}^2
+\end{equation*}
+and for a given $M > 0$, the order $s$ Sobolev ball is $H^s(\Xset, M) = \set{f: \norm{f}_{H^s(\Xset)} \leq M}$.
+The graph Sobolev seminorm $f^T \Lap_{n,r} f$ should be viewed as a random, discrete approximation to the continuum first-order Sobolev seminorm $|f|_{H^1(\Xset)}^2$, and so in our analysis of Laplacian smoothing we will mainly focus on the first-order Sobolev space $H^1(\Xset)$ hereafter.
+
+Much of statistical analysis of nonparametric regression over Sobolev spaces relies on the Sobolev embedding theorem, which establishes that functions in Sobolev spaces also satisfy other, (even more) classical notions of smoothness.
+Concretely, the Sobolev space $H^1(\Xset)$ will continuously embed either into a Holder space $C^{0,q}(\Xset)$ or an $\Leb^p$ space, depending on the dimension $d$.
 \begin{itemize}
-	\item Roughly speaking, Sobolev spaces contain functions $f \in \Leb^p(\Xset)$ with derivatives which themselves belong to $\Leb^p(\Xset)$.
-More formally, for given integers $s$ and $p > 0$, the Sobolev space $W^{s,p}(\Xset)$ consists of all functions $f \in \Leb^p(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ exists and belongs to $\Leb^p(\Xset)$.
-For such functions, the $(s,p)$-Sobolev seminorm $\seminorm{f}_{W^{s,p}(\Xset)}$ and norm $\norm{f}_{W^{s,p}(\Xset)}$ are given by
-	\begin{equation*}
-	\seminorm{f}_{W^{s,p}(\Xset)}^p = \sum_{\abs{\alpha} = s}\int_{\mathcal{X}} \Bigl|\bigl(D^{\alpha}f\bigr)(x)\Bigr|^p \,dx, ~~ \norm{f}_{W^{s,p}(\Xset)}^p = \sum_{k = 0}^{s} \seminorm{f}_{W^{k,p}(\Xset)}^p
-	\end{equation*}
-	and for a given $M > 0$, the $(s,p)$-Sobolev ball is $W^{s,p}(\Xset, M) = \set{f: \norm{f}_{W^{s,p}(\Xset)} \leq M}$.
-In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space, and we adopt the usual convention of writing $H^s(\Xset) = W^{s,2}(\Xset)$ and $H^s(\Xset,M) = W^{s,2}(\Xset,M)$.
-Linear estimators such as $\wh{f}_{\LS}$\footnote{To be explicit, estimators of the form $\wh{f} = HY$ for some matrix $H \in \Reals^{n \times n}$.} can be very accurate over the Hilbert-Sobolev spaces $H^s(\Xset)$, and we will confine our attention to this case hereafter.
-	\item \textcolor{red}{(TODO)}: Introduce the Sobolev embedding theorem, for at least three reasons: (1) to facilitate later discussion of why smoothing splines are ill-posed when $2s < d$; (2) to make explicit why the statistical problem isn't ill-posed when $2s < d$ in the random design setting; (3) to make clear why the testing problem is hopeless when $4s < d$ without imposing additional conditions.
+	\item When $d = 1$, any function $f \in H^1(\Xset)$ satisfies $\|f\|_{H^1(\Xset)} \leq C\|f\|_{C^{0,q}(\Xset)}$ for $q = 1/2.$
+	\item When $d > 2$, any function $f \in H^1(\Xset)$ satisfies $\|f\|_{H^1(\Xset)} \leq C\|f\|_{\Leb^{p}(\Xset)}$ for $p = 2d/(d - 2)$. 
 \end{itemize}
+In the above, the exponents $p$ and $q$ cannot be improved upon.
 
-\paragraph{(2.b) Connecting Laplacian smoothing to smoothing splines.} 
-To better motivate Laplacian smoothing as a tool well-suited for nonparametric regression, we review their connection to one of the most classical methods of nonparametric regression: smoothing splines.
-\begin{itemize}
-	\item Smoothing splines are particularly well-suited for nonparametric regression over Sobolev spaces.
-This is because smoothing splines are penalized least squares estimators where the penalty is itself the Sobolev semi-norm.
-Mathematically, the smoothing spline estimator $\wh{f}_{\SM}$ is defined as the solution to the following optimization problem,
+% AJG commented this out.
+% When $d = 1$, the embedding of $H^1(\Xset)$ into $C^{0,q}(\Xset)$ facilitates a general analysis of (penalized) least estimators over $H^1(\Xset)$, using arguments from empirical process theory.
+However when $d \geq 2$, such arguments fall apart: for example, good bounds on the metric entropy of $H^1(\Xset)$ are not known in this setting.
+More troubling still, least squares estimators over $H^1(\Xset)$ are not even well-posed when $d \geq 2$.
+
+% For these reasons, analysis of least squares estimators for the nonparametric regression problem over Sobolev spaces typically assume $d = 1$---or more generally that $f_0 \in H^s(\Xset)$ for some $2s > d$.
+However, the Laplacian smoother $\wh{f}_{\LS}$ is a special least-squares estimator---it has a nice closed-form solution amenable to a bias-variance decomposition.
+We therefore do not need to rely on empirical process theory, and so we will be able to analyze $\wh{f}_{\LS}$ and $T_{\LS}$ beyond the univariate setting.
+
+As we will see, the Sobolev embedding theorem has rather grave consequences for least-squares estimators over $H^1(\Xset)$ when $d > 1$, rendering such estimators ill-posed.
+However, the Laplacian smoother $\wh{f}_{\LS}$ is a special least-squares estimator.
+It has a nice closed-form solution amenable to a bias-variance decomposition, and so we will be able to analyze $\wh{f}_{\LS}$ and $T_{\LS}$ beyond the univariate setting.
+
+\paragraph{Connecting Laplacian smoothing to smoothing splines.} 
+To better motivate Laplacian smoothing as a tool for nonparametric regression, we connect it to one of the most classical methods of nonparametric regression: smoothing splines.
+
+Smoothing splines are specially designed for regression over Sobolev spaces, as they are penalized least squares estimators where the penalty is itself a Sobolev semi-norm.
+Mathematically, the (first-order) smoothing spline estimator $\wh{f}_{\SM}$ is defined as the solution to the following optimization problem,
 \begin{equation}
 \label{eqn:smoothing_spline}
-	\wt{f} = \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \cdot \seminorm{f}_{H^{1}(\Xset)}^2\biggr\}.
+\wt{f}_{\SM} = \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \cdot \seminorm{f}_{H^{1}(\Xset)}^2\biggr\}.
 \end{equation}
-	In~\eqref{eqn:smoothing_spline}, the minimum can either be taken over all $f \in H^1(\Xset)$ or over all $f \in C^1(\Xset)$; we shall refer to the former as smoothing splines, and the latter as \emph{Holder} smoothing splines.
+In~\eqref{eqn:smoothing_spline}, the minimum can either be taken over all $f \in H^1(\Xset)$ or over all $f \in C^1(\Xset)$; we shall refer to the former as smoothing splines, and the latter as \emph{Lipschitz} smoothing splines.
 
-	\textcolor{red}{(TODO)}: Check the classical spline references, and/or Ryan, for data on (1) what the canonical domain of minimization is for smoothing splines, and (2) whether the representer theorem depends holds for either domain of minimization.
-
-	\item By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can tie the penalty used by the smoothing spline estimator $\wh{f}_{\SM}$ to that used by the Laplacian smoothing estimator $\wh{f}_{\LS}$.
+By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can tie the penalty used by the smoothing spline estimator $\wt{f}_{\SM}$ to that used by the Laplacian smoothing estimator $\wh{f}_{\LS}$.
 Suppose for the moment that $P$ is the uniform distribution over $\Xset$.
-Using the definition~\eqref{eqn:laplace_operator}, integrating by parts, and invoking the aforementioned convergence properties of $\Lap_{n,r}$, we have that for any $f \in C^{2}(\Xset)$ which satisfies appropriate boundary conditions,
+Using the definition~\eqref{eqn:laplace_operator}, integrating by parts, and invoking the aforementioned convergence properties of $\Lap_{n,r}$, we have that for any $f \in C^{2}(\Xset)$ which satisfies Dirichlet boundary conditions (see footnote~\ref{footnote:dirichlet_eigenvalues}),
 \begin{equation}
 \label{eqn:seminorm_convergence}
-	\abs{f}_{H^1(\Xset)} = \int_{\Xset} \bigl(\Delta_P f\bigr)(x) \cdot f(x) \,dx = \lim_{n \to \infty} \frac{1}{n^{2}r^{(d + 2)}}f^T \Lap_{n,r} f.
+\abs{f}_{H^1(\Xset)}^2 = \int_{\Xset} \bigl(\Delta_P f\bigr)(x) \cdot f(x) \,dx = \lim_{n \to \infty} \frac{1}{n^{2}r^{(d + 2)}}f^T \Lap_{n,r} f.
 \end{equation}
-	\textcolor{red}{(TODO)}: Formalize the boundary conditions---I believe either Neumann or Dirichlet boundary conditions should suffice?
-	\item \eqref{eqn:seminorm_convergence} establishes asymptotic equality (up to a scaling factor) between the penalties in~\eqref{eqn:smoothing_spline} and~\eqref{eqn:laplacian_smoothing}, when the distribution $P$ is uniform.
-But it also sheds light on additional feature of Laplacian smoothing, relevant when $P$ is non-uniform: the penalty $f^T \Lap_{n,r} f$ automatically adapts to the design density $p$, in a way that the penalty $|f|_{H^1(\Xset)}$ does not.
-This automatic adaptivity to $P$ turns out to be a formal advantage of Laplacian smoothing, which we will comment on further in Section~\ref{sec:minimax_optimal_laplacian_smoothing}.
-\end{itemize}
 
-\paragraph{(2.c) Minimax-optimality of smoothing splines.} 
-One reason all of this is intriguing is that, as mentioned previously, the smoothing spline estimator $\wh{f}_{\SM}$---and the resultant test statistic $T_{\SM} := \|\wt{f}_{\SM}\|_n^2$--- have strong statistical properties.
+\eqref{eqn:seminorm_convergence} establishes asymptotic equality (up to a scaling factor) between the penalties in~\eqref{eqn:smoothing_spline} and~\eqref{eqn:laplacian_smoothing}, when the distribution $P$ is uniform.
+But it also sheds light on an additional feature of Laplacian smoothing, relevant when $P$ is non-uniform: the penalty $f^T \Lap_{n,r} f$ automatically adapts to the design density $p$, in a way that the penalty $|f|_{H^1(\Xset)}^2$ does not.
+This automatic adaptivity to $P$ turns out to be a formal advantage of Laplacian smoothing relative to smoothing splines, as we will see in Section~\ref{sec:minimax_optimal_laplacian_smoothing}.
+
+\paragraph{Minimax-optimality of smoothing splines.} 
+One reason all of this is intriguing is that, as mentioned previously, the smoothing spline estimator $\wt{f}_{\SM}$---and the resultant test statistic $T_{\SM} := \|\wt{f}_{\SM}\|_n^2$--- have strong statistical properties.
 To be precise, they achieve the well-known minimax rates over Sobolev spaces.
-\begin{itemize}
-	\item  The minimax estimation rate over Sobolev balls, with error measured in $\Leb^2(\Xset)$ norm, is 
+
+Under~\ref{asmp:domain}, the minimax estimation rate over Sobolev balls is (see e.g. \citep{tsybakov2008_book})
 \begin{equation}
 \label{eqn:sobolev_space_estimation_minimax_rate}
-	\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{d/(2 + d)}n^{-2/(2 + d)}~~\textrm{for all $d \geq 1$.}
+\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{2d/(2 + d)}n^{-2/(2 + d)}
 \end{equation}
-
-	\textcolor{red}{(TODO)}: (1) You need a reference for this statement, and you already know one will be tricky to find because of ill-definedness of point evaluation.
-(2) This statement only holds under some condition on the density $p$.
-
-	\item On the other hand, the minimax critical radius over Sobolev balls---that is, the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has non-trivial power over all $H_{\epsilon}^1(\Xset,M) := H^1(\Xset,M) \cap \{f: \norm{f}_{\Leb^2(\Xset)} \geq \epsilon\}$---is much smaller,
+In the goodness-of-fit testing problem, we ask for a test function which can distinguish between the null and alternative hypotheses
+\begin{equation}
+\mathbf{H}_0: f_0 = 0, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{0\}
+\end{equation} 
+where $\mc{F}$ is some class of functions.
+The minimax critical radius is the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{F}_{\epsilon} := \mc{F} \cap \{f: \|f\|_{\Leb^2(\Xset)} \geq \epsilon\}$.
+Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing critical radius over $H^1(\Xset,M)$ is much smaller than the minimax estimation rate:
 \begin{equation}
 \label{eqn:sobolev_space_testing_critical_radius}
-	\epsilon_n^{\star}\Bigl(H^1(\Xset,M)\Bigr):= \inf\Biggl\{\epsilon: \inf_{\phi} \biggl[\Ebb_0[\phi] +  \sup_{f_0 \in H_{\epsilon}^1(\Xset,M)} \Ebb_{f_0}[1 - \phi]\biggr] \leq 2\alpha \Biggr\} \asymp n^{-2/(4 + d)}~~\textrm{for $1 \leq d < 4$,}
+\epsilon^2\Bigl(H^1(\Xset,M)\Bigr):= \inf\Biggl\{\epsilon^2: \inf_{\phi} \biggl[\Ebb_0[\phi] +  \sup_{f_0 \in H_{\epsilon}^1(\Xset,M)} \Ebb_{f_0}[1 - \phi]\biggr] \leq 2\alpha \Biggr\} \asymp M^{2d/(4 + d)}n^{-4/(4 + d)}~~\textrm{for $1 \leq d < 4$;}
 \end{equation}
+see \cite{ingster09} for a proof of this statement, and \cite{ariascastro2018} for a more extended discussion of the minimax paradigm in nonparametric testing.
+When $d \geq 4$ we no longer have a continuous embedding of $H^1(\Xset)$ into $\Leb^4(\Xset)$; the functions in $H^1(\Xset)$ are very irregular, and the minimax rates in this regime are unknown.
 
-	\textcolor{red}{(TODO)}: (1) You need a reference for this statement (Ingster), and (2) you need to explain what happens when $d \geq 4$.
+The estimator $\wt{f}_{\SM}$ and the level-$\alpha$ test $\wt{\phi}_{\SM} := \1\{T_{\SM} \geq \wt{t}_{\alpha}\}$ are known to be minimax rate-optimal---meaning they achieve the rates in~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius}, respectively---when $d = 1$ (see e.g.
+\cite{vandergeer2000} for the estimation result, and \cite{liu2019} for the testing result).
+When $d > 1$, as a consequence of the Sobolev Embedding Theorem\footnote{Or to be more precise, as a consequence of the \emph{tightness} of the Sobolev Embedding Theorem.} the variational problem~\eqref{eqn:smoothing_spline} is not well-posed: as illustrated in e.g.
+\cite{green93}, the criterion is minimized by interpolating the responses $\mathbf{Y}$, and the resulting estimator $\wt{f}_{\SM}$ will be inconsistent.
+By replacing the domain of minimization $H^1(\Xset)$ by $C^1(\Xset)$, we can force the solution to~\eqref{eqn:smoothing_spline} to be less wild, and \cite{birge1993} show that the resulting Lipschitz smoothing spline is not merely consistent but in fact minimax optimal when $d = 2$.\footnote{At least in the estimation problem.
+We could not find either a positive or negative result for the testing problem.} However when $d > 2$, the Holder space $C^1(\Xset)$ is itself too large, and even the Lipschitz smoothing spline has risk provably larger than optimal \cite{birge1993}.
+Previewing things a bit, we will see in Section~\ref{sec:minimax_optimal_laplacian_smoothing} that a similar phenomenon occurs for the Laplacian smoother $\wh{f}$, but interestingly the breakdown point will come when $d > 4$, rather than $d > 1$ or $d > 2$.
 
-	\item The estimator $\wh{f}_{\SM}$ and the level-$\alpha$ test $\phi_{\SM} := \1\{T_{\SM} \geq {t}_{\alpha}\}$ are known to be minimax rate-optimal---meaning they achieve the rates in~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius}, respectively---when $d = 1$.
-When $d > 1$, as a consequence of the Sobolev Embedding Theorem\footnote{Or to be more precise, as a consequence of the \emph{tightness} of the Sobolev Embedding Theorem.} the variational problem~\eqref{eqn:smoothing_spline} is not well-posed: the criterion is minimized by interpolating the responses $\mathbf{Y}$, and the resulting estimator $\wt{f}$ will be inconsistent.
-By replacing the domain of minimization $H^1(\Xset)$ by $C^1(\Xset)$, we can force the solution to~\eqref{eqn:smoothing_spline} to be less wild, and the resulting Holder smoothing spline is not merely consistent but in fact minimax optimal when $d = 2$.\footnote{At least in the estimation problem.
-We could not find either a positive or negative result for the testing problem.} However when $d > 2$, the Holder space $C^1(\Xset)$ is itself too large, and even the Holder smoothing spline has risk provably larger than optimal.
-Previewing things a bit, we will see in Section~\ref{sec:minimax_optimal_laplacian_smoothing} that a similar phenomenon occurs for the Laplacian smoother $\wh{f}$; interestingly, the breakdown point will come when $d > 4$, rather than $d > 1$ or $d > 2$.
+Finally, it is well known that if $\Xset$ satisfies the manifold hypothesis, meaning informally that it has intrinsic dimension $m < d$ (see~\ref{asmp:domain_manifold} in Section~\ref{sec:graph_sobolev_classes} for a formal statement), both~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$ \citep{bickel2007,ingster2000}.
+Encouragingly, the smoothing spline estimator $\wt{f}_{\SM}$ and test $\wt{\phi}_{\SM}$ achieve these faster rates, with one caveat: in order for smoothing splines to properly leverage the manifold hypothesis, the domain $\Xset$ must be known.
 
-	\textcolor{red}{(TODO)}: Include references.
-
-	\item Finally, it is well known that if $\Xset$ satisfies the manifold hypothesis, meaning informally that it has intrinsic dimension $m < d$ (see~\ref{asmp:domain_manifold} in Section~\ref{sec:graph_sobolev_classes} for a formal statement), both~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
-Encouragingly, the smoothing spline estimator $\wt{f}$ and test $\wt{\phi}$ achieve these faster rates, with one caveat: in order for smoothing splines to properly leverage the manifold hypothesis, the domain $\Xset$ must be known.
-\end{itemize}
-
-\paragraph{(2.d) Consequences?} Intuitively, the connections between Laplacian smoothing and smoothing splines give us reason to hope that the strong statistical properties of the latter might be shared by the former.
-Unfortunately, it is not so easy to mathematically tie $\wh{f}$ to $\wt{f}$: for example, note that the convergence in~\eqref{eqn:seminorm_convergence} assumes $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:smoothing_spline} is either $H^1(\Xset)$ or $C^1(\Xset)$.
-Instead, we will take a different approach and analyze $\wh{f}$ by leveraging properties of the graph Sobolev class $H^1(G_{n,r})$, which we now introduce.
+\paragraph{Consequences?} Intuitively, the connections between Laplacian smoothing and smoothing splines give us reason to hope that the strong statistical properties of the latter might be shared by the former.
+Unfortunately, it is not so easy to mathematically tie $\wh{f}_{\LS}$ to $\wt{f}_{\SM}$: for example, note that the convergence in~\eqref{eqn:seminorm_convergence} assumes $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:smoothing_spline} is either $H^1(\Xset)$ or $C^1(\Xset)$.
+Instead, we will take a different approach and analyze $\wh{f}_{\LS}$ by leveraging properties of the graph Sobolev class $H^1(G_{n,r})$, which we now introduce.
 
@@
 
-To understand why graph Sobolev balls play a critical role in the analysis of Laplacian smoothing, let us return for a moment to the continuous-time setting.
+To understand the role graph Sobolev balls play in our analysis of Laplacian smoothing, let us return for a moment to the continuous-time setting.
 Upper bounds on the error of smoothing splines rest on two crucial facts, both of which involve the continuum Laplace operator $\Delta_P$.
-The first is the \textit{a priori} assumption that the Sobolev semi-norm $|f|_{H^1(\Xset)} = \langle f, \Delta_P f \rangle_{\Leb^2(P)} \leq M^2$.\footnote{The equality holding when $P$ is uniform and  only for functions $f$ which satisfy boundary conditions.} The second is an upper bound on the metric entropy of $H^1(\Xset)$, characterized by the eigenvalue decay $\lambda_k(\Delta_P) \asymp k^{2/d}$ (i.e.
+The first is the \textit{a priori} assumption that the Sobolev semi-norm $|f_0|_{H^1(\Xset)}^2 = \langle f_0, \Delta_P f_0 \rangle_{\Leb^2(P)} \leq M^2$.\footnote{The equality holding when $P$ is uniform and  for functions $f_0 \in C^2(\Xset)$ which satisfy Dirichlet conditions.} The second is an upper bound on the metric entropy of $H^1(\Xset,M)$, characterized by the eigenvalue decay $\lambda_k(\Delta_P) \asymp k^{2/d}$ (i.e.
 Weyl's Law).
-In this section we will show that both these statements have analogues in the graph setting:
+In this section we will show that both these statements have analogues in the graph setting.
 \begin{itemize}
-	\item for functions $f$ belonging to $H^1(\Xset,M)$, the evaluations $f \in (f(X_1),\ldots,f(X_n)) \in \Reals^n$ are upper bounded,
+	\item For functions $f$ belonging to $H^1(\Xset,M)$, the evaluations $f \in (f(X_1),\ldots,f(X_n)) \in \Reals^n$ are upper bounded,
 	\begin{equation}
@@
 \end{itemize}
-Both statements will hold with high probability, and when we assume $\Xset$ is a manfiold of dimension $m$, they will hold with $d$ replaced by $m$.
+Both statements will hold with high probability, and when we assume $\Xset$ is a manifold of intrinsic dimension $m$, they will hold with $d$ replaced by $m$.
 These results elucidate why Laplacian smoothing ``works'' as well as its continuous-time counterpart ---in the sense of being statistically rate-optimal over (continuum) Sobolev classes.
 
-\paragraph{(3.a) Graph Sobolev semi-norm.}
-We begin by upper bounding $f_0^T \Lap_{n,r} f_0$ under the assumption $f_0 \in H^1(\Xset,M)$. We will assume the kernel $K$ satisfies condition~\ref{asmp:kernel}.
+\paragraph{Graph Sobolev semi-norm.}
+We begin by upper bounding $f^T \Lap_{n,r} f$ under the assumption $f_0 \in H^1(\Xset,M)$. We will assume the kernel $K$ satisfies condition~\ref{asmp:kernel}.
 \begin{enumerate}[label=(K\arabic*)]
@@
 
-\textbf{(3.b) Graph Laplacian eigenvalues.} The graph eigenvalues $\lambda_k(\Lap_{n,r}) \asymp n r^{d + 2}k^{2/d}$---when $\Xset$ is full-dimensional---or $\lambda_k(\Delta_P) \asymp nr^{m + 2}k^{2/m}$---when $\Xset$ has intrinsic dimension $1 < m < d$.
+\paragraph{Graph Laplacian eigenvalues.} Lemma~\ref{lem:neighborhood_eigenvalue} establishes upper and lower bounds on the graph Laplacian eigenvalues $\lambda_{k}(G_{n,r})$ in terms of the continuum Laplace eigenvalues $\lambda_k(\Delta_P)$.
+These bounds hold with high probability under appropriate regularity conditions on the domain $\Xset$, and for a range of radii $r$.
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{1}
+	\item
+	\label{asmp:domain}
+	The domain $\Xset$ is an open, connected set with Lipschitz boundary.
+\end{enumerate}
+\begin{enumerate}[label=(R\arabic*)]
+	\item
+	\label{asmp:radius1}
+	The radius $r$ satisfies $(\log n/n)^{1/d} \leq r \leq c_1$.
+\end{enumerate}
+\begin{lemma}
+	\label{lem:neighborhood_eigenvalue}
+	Suppose the density $p$, the domain $\Xset$, the kernel $K$ and the radius $r$ satisfy~\ref{asmp:bounded_lipschitz_density}, \ref{asmp:domain}, \ref{asmp:kernel},and~\ref{asmp:radius1} respectively.
+The following statement holds with probability at least $1 - A_1n\exp(-a_1nr^{d})$: for any $\ell \in \mathbb{N}$ such that
+	\begin{equation}
+	\label{eqn:neighborhood_eigenvalue_1}
+	1 - A_1 r \sqrt{\lambda_{\ell}(\Delta_P)} \geq \frac{1}{2}
+	\end{equation}
+	we have that
+	\begin{equation}
+	\label{eqn:eigenvalue_bound}
+	a_1 \lambda_k(G_{n,r}) \leq nr^{d+2} \lambda_k(\Delta_P) \leq A_1 \lambda_k(G_{n,r}),~~\textrm{for all $1 \leq k \leq \ell$}
+	\end{equation}
+\end{lemma}
+
+The scaling~\eqref{eqn:neighborhood_graph_eigenvalue_vague} then follows by Weyl's Law (see Lemma 28 of~\citep{dunlop2020} for a proof that~\ref{asmp:bounded_lipschitz_density} and~\ref{asmp:domain} are sufficient for Weyl's Law to hold.).
+\begin{corollary}
+	\label{cor:neighborhood_eigenvalue}
+	Suppose the density $p$ and the domain $\Xset$ satisfy~\ref{asmp:bounded_lipschitz_density} and~\ref{asmp:domain}; then there exist constants $a_2$ and $A_2$ such that
+	\begin{equation}
+	\label{eqn:weyls_law}
+	a_2k^{2/d} \leq \lambda_k(\Delta_P) \leq A_2k^{2/d}~~\textrm{for all $k \in \mathbb{N}, k > 1$}.
+	\end{equation}
+	Therefore under the conditions of Lemma~\ref{lem:neighborhood_eigenvalue}, we have that
+	\begin{equation}
+	\label{eqn:neighborhood_eigenvalue_2}
+	a_3 \cdot \min\Bigl\{nr^{d + 2}k^{2/d},nr^d\Bigr\} \leq \lambda_k(G_{n,r}) \leq A_3 \cdot \min\Bigl\{nr^{d + 2}k^{2/d}, nr^d\Bigr\}~~\textrm{for all $2 \leq k \leq n$}
+	\end{equation}
+	with probability at least $1 - A_1n\exp(-a_1nr^d)$.
+\end{corollary}
+
+We see that the inequalities in~\eqref{eqn:neighborhood_eigenvalue_2} match the desired scaling~\eqref{eqn:neighborhood_graph_eigenvalue_vague} for all $2 \leq k \leq (1/r)^d$.
+Indeed, if we could choose $r = n^{-1/d}$, we would get the desired scaling for all $2 \leq k \leq n$, but unfortunately choosing $r$ to be so small would violate~\ref{asmp:radius1}.
+Nevertheless, by choosing $r$ to be relatively small while still being in accordance with~\ref{asmp:radius1}, we can guarantee that only some of the largest eigenvalues $\lambda_k(G_{n,r})$ will fail to satisfy~\eqref{eqn:neighborhood_graph_eigenvalue_vague}.
+Fortunately, the behavior of these large eigenvalues will not meaningfully contribute to the error of $\wh{f}_{\LS}$ or $T_{\LS}$, as we shall see.
+
+\paragraph{Manifold hypothesis.}
+
+As mentioned in the introduction, an attractive quality of the neighborhood graph $G_{n,r}$ is that it captures the geometry of $P$ and $\Xset$.
+We now review two manifestations of this phenomenon, from \cite{trillos2019} and \cite{calder2019}: both the graph Sobolev semi-norm and graph Laplacian eigenvalues automatically adapt to the intrinsic dimension of $\Xset$.
+Formally, we make the following assumption on $\Xset$ (in order to directly use the results of~\cite{trillos2019} and~\cite{calder2019}).
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{2}
+	\item 
+	\label{asmp:domain_manifold}
+	The measure $P$ is supported on a compact, connected, smooth submanifold $\Xset$ of $\Reals^d$; the domain $\Xset$ is of fixed dimension $1 \leq m \leq d$ and without boundary.
+Moreover, $P$ has a density $p$ with respect to the volume form of $\Xset$ which is bounded away from $0$ and $\infty$,
+	\begin{equation*}
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty
+	\end{equation*}
+	for all $x \in \Xset$, and the density $p$ is Lipschitz with Lipschitz constant $L_p$.
+\end{enumerate}
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{1}
+	\item 
+	\label{asmp:radius2}
+	The radius $r$ satisfies $(\log n/n)^{1/m} \leq r \leq c_2$.
+\end{enumerate}
+\begin{proposition}[Lemma 5 of~\cite{trillos2019}]
+	\label{prop:manifold_graph_sobolev_seminorm}
+	Assume~\ref{asmp:kernel},~\ref{asmp:domain_manifold} and~\ref{asmp:radius2}. Then for any $f \in H^1(\Xset)$,
+	\begin{equation*}
+	f^T \Lap_{n,r} f \leq \frac{(1 + L_pr)^2}{\delta}n^{2}r^{m + 2} \sigma_K |f|_{H^1(\Xset)}^2
+	\end{equation*}
+	with probability at least $1 - \delta$.
+\end{proposition}
+
+\begin{proposition}[Restatement of Theorem 4 of~\cite{trillos2019}]
+	\label{prop:manifold_neighborhood_eigenvalue}
+	Assume~\ref{asmp:kernel}, ~\ref{asmp:domain_manifold}, and~\ref{asmp:radius2}. It holds that
+	\begin{equation}
+	\label{eqn:manifold_neighborhood_eigenvalue}
+	a_5 \cdot \min\Bigl\{nr^{m + 2}k^{2/m},nr^d\Bigr\} \leq \lambda_k(G_{n,r}) \leq A_5 \cdot \min\Bigl\{nr^{m + 2}k^{2/m}, nr^m\Bigr\}~~\textrm{for all $2 \leq k \leq n$}	
+	\end{equation}
+	with probability at least $1 - A_4n\exp(-a_4nr^m)$.
+\end{proposition}
+Note that~\ref{asmp:domain_manifold} assumes $\Xset$ is without boundary, and so Propositions~\ref{prop:manifold_graph_sobolev_seminorm} and~\ref{prop:manifold_neighborhood_eigenvalue} do not imply Lemmas~\ref{lem:graph_sobolev_seminorm} or~\ref{lem:neighborhood_eigenvalue}.
 
@@
 
-\textbf{(4)} We now leverage the results of Section~\ref{sec:graph_sobolev_classes} to formalize the main conclusions of this work: that Laplacian smoothing methods are minimax rate-optimal.
+We now leverage the results of Section~\ref{sec:graph_sobolev_classes} to formalize the main conclusions of this work: that Laplacian smoothing methods are minimax rate-optimal.
 
-\textbf{(4.a) Estimation error of Laplacian smoothing.} Using \textbf{(3.a)} and \textbf{(3.b)}, we establish upper bounds on the estimation error of Laplacian smoothing.
+For all results in this section, we will assume that the density~$p$ satisfies~\ref{asmp:bounded_lipschitz_density} and the kernel $K$ satisfies~\ref{asmp:kernel}.
+Additionally, within our theorem statements we will use $c$ and $C$ to denote constants which possibly differ from theorem to theorem, and which may depend on $K$, $d$, $p$, and $\Xset$ but do not depend on $f_0$, the radius of the Sobolev ball $M$, or the sample size $n$.
 
-\textbf{(4.b) Testing error of Laplacian smoothing.} Using \textbf{(3.a)} and \textbf{(3.b)}, we establish upper bounds on the testing error of Laplacian smoothing.
+\paragraph{Estimation error of Laplacian smoothing.} 
+Suppose the domain $\Xset$ satisfies~\ref{asmp:domain}.
+Under appropriate conditions, the Laplacian smoothing estimator $\wh{f}_{\LS}$ achieves, up to constants, the minimax risk $n^{-2/(2 + d)}$ over all $f_0 \in H^1(\Xset,M)$.
+To prove this, we will invoke Lemma~\ref{lem:graph_sobolev_seminorm} and Corollary~\ref{cor:neighborhood_eigenvalue} to suitably upper bound the squared-bias and variance of $\wt{f}_{\LS}$, respectively.
+In order to do so, we must restrict the range of the neighborhood graph radius $r$.
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{2}
+	\item 
+	\label{asmp:ls_kernel_radius_estimation}
+	The neighborhood graph radius $r$ satisfies
+	\begin{equation*}
+	\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq r \leq n^{-3/(4 + 2d)} M^{(d - 4)/(4 + 2d)}.
+	\end{equation*}
+\end{enumerate}
+\begin{theorem}
+	\label{thm:laplacian_smoothing_estimation1}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d < 4$ and $f_0 \in H^1(\Xset,M)$.
+Suppose that that the neighborhood graph $G_{n,r}$ is computed with a radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_estimation},  and the Laplacian smoothing estimator $\wh{f}_{\LS}$ with $\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$.
+It holds that
+	\begin{equation*}
+	\Bigl\|\wh{f}_{\LS} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{d/(2 + d)} n^{-2/(2 + d)}
+	\end{equation*}
+	with probability at least $1 - \delta -  A_1n\exp(-a_1nr^d) - \exp(-c M^{d/(2d + 4)} n^{d/(2+d)})$.
+\end{theorem}
+To summarize: when $d = 1,2$ or $3$, the Laplacian smoothing estimator $\wh{f}_{\LS}$ has in-sample mean squared error within a constant factor of the minimax risk, except on a set of small probability.
+Some remarks:
+\begin{itemize}
+	\item When $d = 4$, our analysis results in an upper bound on the mean squared error within a $\log n$ factor of the minimax risk, but when $d \geq 5$ our analysis yields upper bounds that are much worse than the minimax rate.
+This mirrors the conclusions of~\cite{sadhanala16}, who investigate estimation rates of Laplacian smoothing over the $d$-dimensional grid graph.
+In our setting, when $d = 4$ we would set $r \asymp (\log n/n)^{1/4}$---slightly outside the scaling~\ref{asmp:ls_kernel_radius_estimation}---to get the near-optimal rate.
+	\item The range of radii $r$ imposed by~\ref{asmp:ls_kernel_radius_estimation} is  compatible with practice, where by far the most common choice of radius is the connectivity threshold $r \asymp (\log(n)/n)^{1/d}$, chosen to make the graph $G_{n,r}$ as sparse as possible while still being connected.
+	\item The tuning parameter $\rho$ is chosen to balance the (conditional on ${\bf X}$) squared-bias and variance, i.e. chosen as large as possible while still with high probability satisfying
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_bias_variance}
+	\frac{\rho}{n} f_0^T \Lap_{{n,r}} f_0 \lesssim \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k}(G_{n,r}) + 1\bigr)^2}.
+	\end{equation}
+	To balance bias and variance as in~\eqref{eqn:laplacian_smoothing_bias_variance} we require control on all the graph eigenvalues $\lambda_1(G_{n,r}),\ldots,\lambda_n(G_{n,r})$.
+As mentioned in our commentary after Lemma~\ref{lem:neighborhood_eigenvalue}, when $r$ is small enough the large eigenvalues $\lambda_k(G_{n,r})$ contribute negligibly to the variance term.
+The upper bound in~\ref{asmp:radius1} is the largest value of $r$ such that this statement is still true.
+	\item The Laplacian smoothing estimator $\wh{f}_{\LS}$ is defined only at the design points $X_1,\ldots,X_n$ and we therefore measure error using the in-sample mean squared error $\norm{\cdot}_n^2$.
+Of course the in-sample error is a random variable, and our bound is thus a bound in high probability.
+That being said, it is possible to smoothly extend $\wh{f}_{\LS}$ to be defined on all of $\Xset$, or indeed all of $\Reals^d$---for instance, using the \emph{Nystrom extension}---and evaluate the error of the extension in $\Leb^2(\Xset)$ norm.
+Assuming that the extension of our estimators and the regression function $f_0$ are suitably smooth, tools from empirical process theory will guarantee that the $\Leb^2(\Xset)$ error is not too much greater than the in-sample error, but we do not further pursue the details here.
+\end{itemize}
 
-\quad \textbf{(4.b.i) Low-smoothness regime.} Unlike in the estimation setting, the nonparametric goodness-of-fit testing problem itself exhibits an elbow when $d = 4$ and $f_0 \in \Leb^4(\Xset)$.
-The naive statistic $\frac{1}{n}\|Y\|_2^2$ is now optimal.
-By tuning Laplacian smoothing to the interpolation limit---i.e.
-setting $\rho = 0$---we recover this naive statistic, and therefore Laplacian smoothing is optimal in this sense (although really, no smoothing is being done.)
+\paragraph{Testing error of Laplacian smoothing.}
+Let us define a test using the statistic $T_{\LS}$. For $b \geq 1$, define the threshold $\wh{t}_b$ to be
+\begin{equation*}
+\wh{t}_{b} := \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k(G_{n,r}) + 1\bigr)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k(G_{n,r}) + 1\bigr)^4}},
+\end{equation*}
+and the Laplacian smoothing test $\wh{\phi}_{\LS} := \1\bigl\{T_{\LS} \leq \wh{t}_b\bigr\}$.
+(Clearly the test $\wh{\phi}_{\LS}$ depends on $b$, but we suppress this notationally.) The number $b \geq 1$ is selected by the user, with the choice made based on the tolerated level of Type I and Type II error.
+In fact, the threshold $\wh{t}_b$ is precisely the right choice to control the Type I error of $\wh{\phi}_{\LS}$; as we show in Lemma~\ref{lem:ls_fixed_graph_testing} in the supplementary material,
+\begin{equation*}
+\Ebb_0\Bigl[\wh{\phi}_{\LS}\Bigr] \leq \frac{1}{b^2}.
+\end{equation*}
 
-\textbf{(4.c) Comparing with smoothing splines.} While we motivated graph Laplacians by suggesting they might replicate the statistical properties of smoothing splines, we have in fact managed to show something stronger.
-In particular:
+Theorem~\ref{thm:laplacian_smoothing_testing} establishes that this test also has small Type II error, uniformly over all $f_0$ separated from $0$ by at least the critical radius $\epsilon\bigl(H^1(\Xset,M)\bigr)$ given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
+For this to hold, we will require a tight range of scalings for $r$.
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{3}
+	\item 
+	\label{asmp:ls_kernel_radius_testing}
+	The neighborhood graph radius $r$ satisfies
+	\begin{equation*}
+	\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq r \leq M^{(16 - d)/(8 + 2d)}n^{(d - 20)/(32 + 8d)}.
+	\end{equation*}
+\end{enumerate}
 
-\quad \textbf{(4.c.i)} When $d \geq 2$, there is no Sobolev embedding of $H^1(\Xset) \subseteq C^s(\Xset)$ for any $s > 0$. Therefore the smoothing spline estimator is ill-posed.
+\begin{theorem}
+	\label{thm:laplacian_smoothing_testing}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d < 4$ and $f_0 \in H^1(\Xset,M)$ for $M \leq n^{(4 - d)/(4 + d)}$.
+Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}, and the Laplacian smoothing test $\wh{\phi}_{\LS}$ with $\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$, and $b \geq 1$.
+For all $f_0$ such that
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_testing}
+	\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2d/(4 + d)} n^{-4/(4 + d)}
+	\end{equation} 
+	the Type II error is upper bounded:
+	\begin{equation*}
+	\Ebb_{f_0}\Bigl[1 - \wh{\phi}_{\LS} \Bigr] \leq C\biggl(\frac{1}{b}\Bigl[1 + n^{-d(4 + d)}M^{-2d/(4 + d)}\Bigr] + A_1n\exp\bigl(-a_1nr^d\bigr)\biggr).
+	\end{equation*}
+\end{theorem}
 
-\quad \textbf{(4.c.ii)} Additionally, the metric entropy of $C^1(\Xset)$ is too large to obtain optimal rates for \textcolor{red}{Holder-smoothing splines} when $d \geq 2$.
-Thus, we see a gap emerge between the optimality properties of the discrete-time (Laplacian smoothing) and continuous-time (smoothing splines) estimators, in favor of the latter.
+A few remarks:
+\begin{itemize}
+	\item As mentioned in Section~\ref{sec:minimax_optimal_regression_sobolev_spaces}, when $d \geq 4$ the Sobolev balls $H^1(\Xset,M)$ include quite irregular functions $f \not\in \Leb^4(\Xset)$.
+Proving tight lower bounds over such classes is non-trivial, and to the best of our knowledge such an analysis remains outstanding.
+On the other hand, if we explicitly assume that $f \in \Leb^4(\Xset,M)$, then \cite{guerre02} show that the testing problem is characterized by the dimension-free lower bound $\epsilon^{2}(\Leb^4(\Xset,M)) \gtrsim n^{-1/2}$.
+Moreover, by training $\wh{f}_{\LS}$ to the interpolation limit, i.e.
+setting $\rho = 0$, the test $\wh{\phi}_{\LS}$ achieves (up to constants) this lower bound.
+That is,
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_testing_low_smoothness}
+	E_{f_0}\Bigl[1 - \wh{\phi}_{\LS}\Bigr] \leq \frac{C(1 + M^4)}{b^2}
+	\end{equation}
+	for any $f_0 \in \Leb^4(\Xset,M)$ for which $\|f_0\|_{\Leb^2(\Xset)}^2 \geq C b^2n^{-1/2}$. 
+	\item To compute the threshold $\wh{t}_b$, one must know each of the eigenvalues $\lambda_1(G_{n,r}),\ldots,\lambda_n(G_{n,r})$.
+Computing all $n$ of these eigenvalues is far more expensive than computing $T_{\LS}$.
+That being said, in practice we would not recommend using $\wh{t}_b$ anyway: the threshold is justified by asymptotic theory, but in practice $n$ is always finite, and using such thresholds can result in substantial loss of efficiency.
+We would instead make the standard recommendation to calibrate via permutation.
+\end{itemize}
+It is worth circling back to our comparison of Laplacian smoothing methods to smoothing splines.
+While we motivated Laplacian smoothing methods in part by suggesting they might replicate the statistical properties of smoothing splines, we have in fact managed to show something stronger.
+Theorems~\ref{thm:laplacian_smoothing_estimation1} and~\ref{thm:laplacian_smoothing_testing} imply that Laplacian smoothing methods are minimax optimal for both estimation and testing in dimensions $d = 1,2$ and $3$, and are nearly minimax-optimal for the estimation problem when $d = 4$.
+On the other hand, as covered in Section~\ref{sec:minimax_optimal_regression_sobolev_spaces} smoothing spline methods are optimal only when $d = 1$ or, for Lipschitz smoothing splines, $d = 2$.
+Thus, we see a somewhat surprising gap emerge between the optimality properties of the discrete-time (Laplacian smoothing) and continuous-time (smoothing spline) approaches, in favor of the former.
 
-\quad \textbf{(4.c.iii)} On the other hand, when $d \geq 4$, we no longer get the known optimal rates in the estimation problem.
-When $d = 4$, our rates are suboptimal by a factor of $\log n$.
-When $d > 5$ they are very suboptimal.
-We believe this is related to classical results regarding the entropy of low-smoothness Holder classes, i.e.
-the results used to justify \textbf{(4.c.ii)}.
-The interesting point is that the parameter spaces $H^1(\Xset)$, $C^1(\Xset)$ and $H^1(G_{n,r})$ become ``too large'' for different values of $d$.
+\paragraph{Manifold adaptivity.}
 
-\quad \textbf{(4.c.iv)} \textcolor{red}{(TODO)}: Translate these results into error in $\Leb^2(P)$ error, if/when possible.
+In Section~\ref{sec:graph_sobolev_classes}, we reviewed that when $\Xset$ is a submanifold of $\Rd$, the graph Laplacian automatically adapts to the geometry---and in particular, the intrinsic dimension $m$---of $\Xset$.
+This translates into improved upper bounds on the estimation and testing error of Laplacian smoothing; the upper bounds now depend on $m$ instead of $d$.
 
-\textbf{(4.d) Manifold adaptivity.} In Section~\ref{sec:graph_sobolev_classes}, we established that when $\Xset$ is an $m$-dimensional submanifold of $\Reals^d$, Laplacian smoothing approaches automatically ``feel'' the intrinsic dimension $m$ of $\Xset$.
-We now show that this translates into improved upper bounds on estimating and testing error.
-
-\quad \textbf{(4.d.i)} This confirms the statement in \textbf{(4)} that Laplacian smoothing automatically adapts to the geometry of $\Xset$ and $p$.
-
-\quad \textbf{(4.d.ii)} By contrast, smoothing splines must be specially tailored to work in this setting.
-
-\textbf{(4.e) Higher-order smoothness classes.} Laplacian smoothing can also be adapted to take advantage of additional assumed regularity on $f_0$, i.e.
-$f_0 \in H^s(\mc{\Xset})$ for $s > 1$.
-In the very special case of $s = 2$ and $d = 2$, we can show that this adapted estimator achieves the sharper minimax rates over these classes, but the general story--for all combinations of $s$ and $d$---remains beyond our reach.
+\begin{theorem}
+	\label{thm:laplacian_smoothing_manifold}
+	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
+Suppose that $\Xset$ satisfies~\ref{asmp:domain_manifold} for $m < 4$, and that $f_0 \in H^1(\Xset,M)$.
+For appropriate choices of tuning parameters $r$, $\rho$, and $b \geq 1$, each of the following statements holds.
+	\begin{itemize}
+		\item The Laplacian smoothing estimator satisfies
+		\begin{equation*}
+		\frac{1}{n}\bigl\|\wh{f}_{\LS} - f_0\bigr\|_2^2 \leq \frac{C}{\delta}M^{m/(2 + m)}n^{-2/(2 + m)}
+		\end{equation*}
+		with probability at least $1 - \delta - A_4n\exp(-a_4nr^m) - \exp(-c M^{m/(2m + 4)} n^{m/(2+m)})$.
+		\item Suppose $M \leq n^{(4 - m)/(4 + m)}$. For all $f_0$ such that
+		\begin{equation*}
+		\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2m/(4 + m)} n^{-4/(4 + m)}
+		\end{equation*} 
+		the type II error of $\wh{\phi}$ is upper bounded:
+		\begin{equation*}
+		\Ebb_{f_0}\Bigl[1 - \wh{\phi}_{\LS} \Bigr] \leq C\biggl(\frac{1}{b}\Bigl[1 + n^{-m(4 + m)}M^{-2m/(4 + m)}\Bigr] + A_4n\exp\bigl(-a_4nr^m\bigr)\biggr).
+		\end{equation*}
+	\end{itemize}
+\end{theorem}
+In our supplementary material, we specify the right choices of $\rho$, as well as the permissible ranges of $r$.
+For now, we merely point out that they will depend only on the intrinsic dimension $m$, and not on any other properties of the manifold $\Xset$.
+In all other respects, $\wh{f}_{\LS}$ and $T_{\LS}$ are computed without any knowledge of $\Xset$.
+By contrast, the penalty in~\eqref{eqn:smoothing_spline} must be specially tailored to $\Xset$ in order for $\wt{f}_{\SM}$ and $T_{\SM}$ to achieve the minimax estimation and testing rates.
+This reveals another advantage of Laplacian smoothing as opposed to smoothing splines.
 
 \section{Simulations}
-
-\textbf{(5)} Empirically, we demonstrate that the risk for Laplacian smoothing is comparable to that of smoothing splines---in the full dimensional case with uniformly random design.
-When the design is sampled from a density concentrated near a manifold, Laplacian smoothing outperforms (out of the box) smoothing splines.
-
-\section{Discussion}
+\label{sec:simulations}
 
@@
 
-\clearpage
+\section{Discussion}
+\label{sec:discussion}
 
-\appendix
+\textcolor{red}{(TODO)}
 
-\clearpage
-
-Many of our results follow the same general two-part proof strategy.
-First, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to functionals of the graph $G$.
-We then analyze the behavior of these functionals with respect to the particular neighborhood graph $G_{n,r}$ and give high probability (upper or lower) bounds on these functionals.
-It is in this second step where we invoke our various assumptions on the distribution $P$ and regression function $f_0$.
-
-\section{Graph-dependent error bounds}
-\label{sec:fixed_graph_error_bounds}
-
-\section{Neighborhood graph Sobolev semi-norm}
-\label{sec:graph_sobolev_seminorm}
-
-\section{Bounds on neighborhood graph eigenvalues}
-\label{sec:graph_eigenvalues}
-
-\section{Bounds on the empirical norm}
-\label{sec:empirical_norm}
-
-\section{Proof of theorems}
+A few discussion items:
+\begin{itemize}
+	\item \textbf{Higher-order smoothness classes.} Laplacian smoothing can also be adapted to take advantage of additional assumed regularity on $f_0$, i.e.
+$f_0 \in H^s(\mc{\Xset})$ for $s > 1$.
+In the very special case of $s = 2$ and $d = 2$, we can show that this adapted estimator achieves the sharper minimax rates over these classes, but the general story--for all combinations of $s$ and $d$---remains beyond our reach.
+	\item \textbf{Possible extensions}. Methodologically: kNN graphs and different normalizations of the Laplacian. Theoretically: relaxing~\ref{asmp:bounded_lipschitz_density}. 
+\end{itemize}
 

### 2021-02-11 notes/manuscripts/laplacian_smoothing/AISTATS/archive/outline.tex
Tampered with directory structure.

--- 2021-02-11 notes/manuscripts/laplacian_smoothing/AISTATS/archive/minimax_optimal_laplacian_smoothing.tex
@@
 
-\begin{itemize}
-	\item In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Rd$, and
-	\begin{equation}
-	\label{eqn:random_design_regression}
-	Y_i = f_0(X_i) + \varepsilon_i
-	\end{equation}
-	with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise.
-Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present.
-	\item In graph-based nonparametric regression, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$.
-The neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$.
-The $n \times n$ weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $\mathbf{W}_{ij}$ are given by
-	\begin{equation*}
-	\label{eqn:neighborhood_graph}
-	{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr).
-	\end{equation*}
-	Then the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$, and the graph Laplacian can be written as
-	\begin{equation}
-	\label{eqn:graph_Laplacian}
-	\Lap_{n,r} = \bf{D} - \bf{W}
-	\end{equation}
-	\item The Laplacian smoothing estimator $\wh{f}_{\LS}$ \citep{smola2003} is a penalized least squares estimator, given by
-	\begin{equation}
-	\label{eqn:laplacian_smoothing}
-	\wh{f}_{\LS} := \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_{n,r} f \biggr\}
-	\end{equation}
-	where $\rho > 0$ acts a tuning parameter on the penalty $f^T \Lap_{{n,r}} f$.
-	Assuming~\eqref{eqn:laplacian_smoothing} is a reasonable estimator of $f_0$, the squared empirical norm
-	\begin{equation}
-	T_{\LS} = \frac{1}{n}\Bigl\|\wt{f}_{\LS}\Bigr\|_2^2 \label{eqn:laplacian_smoothing_test}
-	\end{equation}
-	is in turn a reasonable statistic to assess whether or not $f_0 = 0$. 
-\end{itemize}
-
-\textbf{(2.a)} Laplacian smoothing has many advantages compared to other nonparametric regression methods. For instance:
+\textbf{(1.a) Motivation.} Laplacian smoothing has many advantages compared to other nonparametric regression methods. For instance:
 
@@
 
-\textbf{(2.b)} For this reason a considerable body of work has emerged analyzing the consistency of the graph Laplacian $\Lap_{{n,r}}$ as $n \to \infty$ and $r(n) \to 0$.
+\textbf{(1.b) Related work.} For this reason a considerable body of work has emerged analyzing the consistency of the graph Laplacian $\Lap_{{n,r}}$ as $n \to \infty$ and $r(n) \to 0$.
 This is meant in various senses:
@@
 
-\textbf{(2.c)} Our main contributions lie in this latter direction, and they can all be summarized as follows: Laplacian smoothing methods are minimax optimal over Sobolev spaces.
+\textbf{(1.c) Our contributions.} Our main contributions lie in this latter direction, and they can all be summarized as follows: Laplacian smoothing methods are minimax optimal over Sobolev spaces.
 In more detail, we will show that when $f_0$ belongs to the Sobolev ball $H^1(\Xset;M)$:
@@
 
-\paragraph{Organization.}
+\textbf{(1.d) Organization.}
 
-List out what is to come.
+\textbf{(1.e) Notation.}
 
@@
 
-Before we get to our main results, we briefly review minimax optimal estimation and testing rates over Sobolev classes.
+\textbf{(2)} Before we get to our main results, we briefly review minimax optimal estimation and testing rates over Sobolev classes.
 We also single out a notable method which achieves these rates: \emph{smoothing splines}.
@@
 
-\subsection{Sobolev spaces}
-\label{subsec:sobolev_spaces}
+\textbf{(2.a) Sobolev spaces.} Sobolev spaces are classical function spaces, well-suited for linear nonparametric regression methods.
 
-We start by briefly introducing Sobolev spaces, emphasizing only the material relevant to our statistical context.
-Roughly speaking, Sobolev spaces contain functions $f \in \Leb^p(\Xset)$ with derivatives which themselves belong to $\Leb^p(\Xset)$.
-More formally, for given integers $s$ and $p > 0$, the Sobolev space $W^{s,p}(\Xset)$ consists of all functions $f \in \Leb^p(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ exists and belongs to $\Leb^p(\Xset)$.
-For such functions, the $(s,p)$-Sobolev seminorm $\seminorm{f}_{W^{s,p}(\Xset)}$ and norm $\norm{f}_{W^{s,p}(\Xset)}$ are given by
-\begin{equation*}
-\seminorm{f}_{W^{s,p}(\Xset)}^p = \sum_{\abs{\alpha} = s}\int_{\mathcal{X}} \Bigl|\bigl(D^{\alpha}f\bigr)(x)\Bigr|^p \,dx, ~~ \norm{f}_{W^{s,p}(\Xset)}^p = \sum_{k = 0}^{s} \seminorm{f}_{W^{k,p}(\Xset)}^p
-\end{equation*}
-and for a given $M > 0$, the $(s,p)$-Sobolev ball is $W^{s,p}(\Xset, M) = \set{f: \norm{f}_{W^{s,p}(\Xset)} \leq M}$.
-In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space, and we adopt the usual convention of writing $H^s(\Xset) = W^{s,2}(\Xset)$ and $H^s(\Xset,M) = W^{s,2}(\Xset,M)$; we will confine our attention to this case hereafter.
-
-\textbf{(4)} In fact, an alternative way to motivate Laplacian smoothing nonparametric regression methods---as opposed to the advantages mentioned in \textbf{(2.a)}---comes from viewing them as discrete-time versions of smoothing splines.
-This is made more rigorous by the consistency results of \textbf{(2.b)}.
+\textbf{(2.b) Smoothing splines, and their connection to Laplacian smoothing.} One notable such method is smoothing splines, to which we give special attention because of their close connection to Laplacian smoothing.
+In fact, an alternative way to motivate Laplacian smoothing nonparametric regression methods---as opposed to the advantages mentioned in \textbf{(1.a)}---comes from viewing them as discrete-time versions of smoothing splines.
+This is made more rigorous by the consistency results of \textbf{(1.b)}.
 Viewed in this light, an additional advantage of Laplacian smoothing emerges: it automatically adapts to the geometry of $\Xset$ and $p$.
 
-\textbf{(5)} Smoothing splines are known to have strong minimax optimal properties. Reasoning by analogy, we might hope that Laplacian smoothing would share these properties.
+\textbf{(2.c) Optimality of smoothing splines.} Smoothing splines are known to have strong minimax optimal properties when $d = 1$.
+
+\textbf{(2.d) Consequences?} Reasoning by analogy, we might hope that Laplacian smoothing would share these properties.
 
 \section{Graph Sobolev classes}
+\label{sec:graph_sobolev_classes}
 
-\textbf{(6)} The minimax optimality of smoothing splines rests on two crucial facts: one, the a priori assumption that $|f|_{H^1(\Xset)}$ is bounded; two, an upper bound on the metric entropy of $H^1(\Xset)$, characterized by the decay of its eigenvalues.
-\textcolor{red}{(TODO)}:(I know this isn't well-stated, and I need to clean it up.)  We now show that qualitatively similar statements hold with respect to
+\textbf{(3)} The minimax optimality of smoothing splines rests on two crucial facts involving the continuum Laplace operator $\Delta_P$: one, the \textit{a priori} assumption that the Sobolev semi-norm $\langle f, \Delta_P f \rangle_{\Leb^2(P)}$ is bounded; two, an upper bound on the metric entropy of $H^1(\Xset)$, characterized by the eigenvalue decay $\lambda_k(\Delta_P) \sim k^{2/d}$---when $\Xset$ is full-dimensional--- or $\lambda_k(\Delta_P) \sim k^{2/m}$---when $\Xset$ has intrinsic dimension $1 < m < d$ (i.e.
+Weyl's Law).
+We now show that equivalent statements hold with respect to $\Lap_{n,r}$.
 
-\quad \textbf{(6.1)} the graph Sobolev semi-norm $f_0^T \Lap_{n,r} f_0$ and,
+\textbf{(3.a) Graph Sobolev semi-norm.} For functions $f \in H^1(\Xset, M)$, the graph Sobolev semi-norm $f_0^T \Lap_{n,r} f_0 \lesssim M^2 r^{d + 2}n^2$---when $\Xset$ is full-dimensional---or $f_0^T \Lap_{n,r} f_0 \lesssim M^2 r^{m + 2}n^2$---when $\Xset$ has intrinsic dimension $1 < m < d$.
 
-\quad \textbf{(6.2)} the graph eigenvalues $\lambda_k(\Lap_{n,r})$.
+\textbf{(3.b) Graph Laplacian eigenvalues.} The graph eigenvalues $\lambda_k(\Lap_{n,r}) \asymp n r^{d + 2}k^{2/d}$---when $\Xset$ is full-dimensional---or $\lambda_k(\Delta_P) \asymp nr^{m + 2}k^{2/m}$---when $\Xset$ has intrinsic dimension $1 < m < d$.
 
@@
 
-\textbf{(7)} Using \textbf{(6.1)} and \textbf{(6.2)}, we establish upper bounds on the estimation and testing error of Laplacian smoothing, which confirm that Laplacian smoothing is asymptotically minimax rate-optimal over the Sobolev class $H^1(\Xset)$.
+\textbf{(4)} We now leverage the results of Section~\ref{sec:graph_sobolev_classes} to formalize the main conclusions of this work: that Laplacian smoothing methods are minimax rate-optimal.
 
-\textbf{(8)} While we motivated graph Laplacians by suggesting they might replicate the statistical properties of smoothing splines, we have in fact managed to show something stronger. In particular:
+\textbf{(4.a) Estimation error of Laplacian smoothing.} Using \textbf{(3.a)} and \textbf{(3.b)}, we establish upper bounds on the estimation error of Laplacian smoothing.
 
-\quad \textbf{(8.1)} When $d \geq 2$, there is no Sobolev embedding of $H^1(\Xset) \subseteq C^s(\Xset)$ for any $s > 0$. Therefore the smoothing spline estimator is ill-posed.
+\textbf{(4.b) Testing error of Laplacian smoothing.} Using \textbf{(3.a)} and \textbf{(3.b)}, we establish upper bounds on the testing error of Laplacian smoothing.
 
-\quad \textbf{(8.2)} Additionally, the metric entropy of $C^1(\Xset)$ is too large to obtain optimal rates for \textcolor{red}{Holder-smoothing splines} when $d \geq 2$.
+\quad \textbf{(4.b.i) Low-smoothness regime.} Unlike in the estimation setting, the nonparametric goodness-of-fit testing problem itself exhibits an elbow when $d = 4$ and $f_0 \in \Leb^4(\Xset)$.
+The naive statistic $\frac{1}{n}\|Y\|_2^2$ is now optimal.
+By tuning Laplacian smoothing to the interpolation limit---i.e.
+setting $\rho = 0$---we recover this naive statistic, and therefore Laplacian smoothing is optimal in this sense (although really, no smoothing is being done.)
+
+\textbf{(4.c) Comparing with smoothing splines.} While we motivated graph Laplacians by suggesting they might replicate the statistical properties of smoothing splines, we have in fact managed to show something stronger.
+In particular:
+
+\quad \textbf{(4.c.i)} When $d \geq 2$, there is no Sobolev embedding of $H^1(\Xset) \subseteq C^s(\Xset)$ for any $s > 0$. Therefore the smoothing spline estimator is ill-posed.
+
+\quad \textbf{(4.c.ii)} Additionally, the metric entropy of $C^1(\Xset)$ is too large to obtain optimal rates for \textcolor{red}{Holder-smoothing splines} when $d \geq 2$.
 Thus, we see a gap emerge between the optimality properties of the discrete-time (Laplacian smoothing) and continuous-time (smoothing splines) estimators, in favor of the latter.
 
-\quad \textbf{(8.3)} On the other hand, when $d \geq 4$, we no longer get the known optimal rates in the estimation problem.
+\quad \textbf{(4.c.iii)} On the other hand, when $d \geq 4$, we no longer get the known optimal rates in the estimation problem.
 When $d = 4$, our rates are suboptimal by a factor of $\log n$.
@@
 We believe this is related to classical results regarding the entropy of low-smoothness Holder classes, i.e.
-the results used to justify \textbf{(8.2)}.
-The interesting point is that the parameter spaces $C^1(\Xset)$ and $H^1(G_{n,r})$ become ``too large'' for different values of $d$.
+the results used to justify \textbf{(4.c.ii)}.
+The interesting point is that the parameter spaces $H^1(\Xset)$, $C^1(\Xset)$ and $H^1(G_{n,r})$ become ``too large'' for different values of $d$.
 
-\quad \textbf{(8.4)} \textcolor{red}{(TODO)}: Translate these results into error in $\Leb^2(P)$ error, if possible.
+\quad \textbf{(4.c.iv)} \textcolor{red}{(TODO)}: Translate these results into error in $\Leb^2(P)$ error, if/when possible.
 
-\subsection{Manifold adaptivity}
+\textbf{(4.d) Manifold adaptivity.} In Section~\ref{sec:graph_sobolev_classes}, we established that when $\Xset$ is an $m$-dimensional submanifold of $\Reals^d$, Laplacian smoothing approaches automatically ``feel'' the intrinsic dimension $m$ of $\Xset$.
+We now show that this translates into improved upper bounds on estimating and testing error.
 
-\quad \textbf{(9)} We now show that when $\Xset$ is an $m$-dimensional submanifold of $\Reals^d$, Laplacian smoothing approaches automatically ``feel'' the intrinsic dimension $m$ of $\Xset$, which translate into improved upper bounds on estimating and testing error.
-This confirms the statement in \textbf{(4)} that Laplacian smoothing automatically adapts to the geometry of $\Xset$ and $p$.
+\quad \textbf{(4.d.i)} This confirms the statement in \textbf{(4)} that Laplacian smoothing automatically adapts to the geometry of $\Xset$ and $p$.
 
-\subsection{Higher-order smoothness classes}
+\quad \textbf{(4.d.ii)} By contrast, smoothing splines must be specially tailored to work in this setting.
 
-\quad \textbf{(10)} Analogously to smoothing splines, Laplacian smoothing can also be adapted to take advantage of additional assumed regularity on $f_0$, i.e.
+\textbf{(4.e) Higher-order smoothness classes.} Laplacian smoothing can also be adapted to take advantage of additional assumed regularity on $f_0$, i.e.
 $f_0 \in H^s(\mc{\Xset})$ for $s > 1$.
-In very limited circumstances, we can show that this adapted estimator achieves the sharper minimax rates over these classes.
-However, the general story--for all combinations of $s$ and $d$---remains beyond our reach.
+In the very special case of $s = 2$ and $d = 2$, we can show that this adapted estimator achieves the sharper minimax rates over these classes, but the general story--for all combinations of $s$ and $d$---remains beyond our reach.
 
@@
 
-\quad \textbf{(11)} Empirically, we demonstrate that the risk for Laplacian smoothing is comparable to that of smoothing splines, in the full dimensional case.
-Although one cannot carry out a comparison in the case of an unknown manifold --- since the smoothing spline estimator must be changed in a non-obvious manner --- we do show that when the density $p$ is concentrated close to a manifold and $f_0$ is smooth where $p$ is peaked, we have improved performance for $\wh{f}_{\LS}$ compared to $\wh{f}_{\mathrm{SM}}$.
+\quad \textbf{(5)} Empirically, we demonstrate that the risk for Laplacian smoothing is comparable to that of smoothing splines---in the full dimensional case with uniformly random design.
+When the design is sampled from a density concentrated near a manifold, Laplacian smoothing outperforms (out of the box) smoothing splines.
 

### 2021-02-11 notes/manuscripts/laplacian_smoothing/AISTATS/figures.tex
Tampered with directory structure.

\documentclass[twoside]{article}

\usepackage{aistats2021}
\usepackage{graphicx}
% If your paper is accepted, change the options for the package
% aistats2021 as follows:
%
%\usepackage[accepted]{aistats2021}
%
% This option will print headings for the title of your paper and
% headings for the authors names, plus a copyright note at the end of
% the first column of the first page.

% If you set papersize explicitly, activate the following three lines:
%\special{papersize = 8.5in, 11in}
%\setlength{\pdfpageheight}{11in}
%\setlength{\pdfpagewidth}{8.5in}

% If you use natbib package, activate the following three lines:
%\usepackage[round]{natbib}
%\renewcommand{\bibname}{References}
%\renewcommand{\bibsection}{\subsubsection*{\bibname}}

% If you use BibTeX in apalike style, activate the following line:
%\bibliographystyle{apalike}

\begin{document}

% If your paper is accepted and the title of your paper is very long,
% the style will print as headings an error message. Use the following
% command to supply a shorter title of your paper so that it can be
% used as headings.
%
%\runningtitle{I use this title instead because the last one was very long}

% If your paper is accepted and the number of authors is large, the
% style will print as headings an error message. Use the following
% command to supply a shorter version of the authors names so that
% they can be used as headings (for example, use only the surnames)
%
%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}

\twocolumn[

\aistatstitle{Instructions for Paper Submissions to AISTATS 2021}

\aistatsauthor{ Author 1 \And Author 2 \And  Author 3 }

\aistatsaddress{ Institution 1 \And  Institution 2 \And Institution 3 } ]

\begin{abstract}
  The Abstract paragraph should be indented 0.25 inch (1.5 picas) on
  both left and right-hand margins. Use 10~point type, with a vertical
  spacing of 11~points. The \textbf{Abstract} heading must be centered,
  bold, and in point size 12. Two line spaces precede the
  Abstract. The Abstract must be limited to one paragraph.
\end{abstract}

\section{GENERAL FORMATTING INSTRUCTIONS}

The camera-ready versions of the accepted papers are 8 pages,
plus any additional pages needed for references.

Papers are in 2 columns with the overall line width of 6.75~inches (41~picas).
Each column is 3.25~inches wide (19.5~picas).  The space
between the columns is .25~inches wide (1.5~picas).  The left margin is 0.88~inches (5.28~picas).
Use 10~point type with a vertical spacing of
11~points. Please use US Letter size paper instead of A4.

Paper title is 16~point, caps/lc, bold, centered between 2~horizontal rules.
Top rule is 4~points thick and bottom rule is 1~point thick.
Allow 1/4~inch space above and below title to rules.

Author descriptions are center-justified, initial caps.  The lead
author is to be listed first (left-most), and the Co-authors are set
to follow.  If up to three authors, use a single row of author
descriptions, each one center-justified, and all set side by side;
with more authors or unusually long names or institutions, use more
rows.

Use one-half line space between paragraphs, with no indent.

\section{FIRST LEVEL HEADINGS}

First level headings are all caps, flush left, bold, and in point size
12. Use one line space before the first level heading and one-half line space
after the first level heading.

\subsection{Second Level Heading}

Second level headings are initial caps, flush left, bold, and in point
size 10. Use one line space before the second level heading and one-half line
space after the second level heading.

\subsubsection{Third Level Heading}

Third level headings are flush left, initial caps, bold, and in point
size 10. Use one line space before the third level heading and one-half line
space after the third level heading.

\paragraph{Fourth Level Heading}

Fourth level headings must be flush left, initial caps, bold, and
Roman type.  Use one line space before the fourth level heading, and
place the section text immediately after the heading with no line
break, but an 11 point horizontal space.

%%%
\subsection{Citations, Figure, References}

\subsubsection{Citations in Text}

Citations within the text should include the author's last name and
year, e.g., (Cheesman, 1985). 
%Apart from including the author's last name and year, citations can follow any style, as long as the style is consistent throughout the paper.  
Be sure that the sentence reads
correctly if the citation is deleted: e.g., instead of ``As described
by (Cheesman, 1985), we first frobulate the widgets,'' write ``As
described by Cheesman (1985), we first frobulate the widgets.''

The references listed at the end of the paper can follow any style as long as it is used consistently.

%Be sure to avoid
%accidentally disclosing author identities through citations.

\subsubsection{Footnotes}

Indicate footnotes with a number\footnote{Sample of the first
  footnote.} in the text. Use 8 point type for footnotes. Place the
footnotes at the bottom of the column in which their markers appear,
continuing to the next column if required. Precede the footnote
section of a column with a 0.5 point horizontal rule 1~inch (6~picas)
long.\footnote{Sample of the second footnote.}

\subsubsection{Figures}

All artwork must be centered, neat, clean, and legible.  All lines
should be very dark for purposes of reproduction, and art work should
not be hand-drawn.  Figures may appear at the top of a column, at the
top of a page spanning multiple columns, inline within a column, or
with text wrapped around them, but the figure number and caption
always appear immediately below the figure.  Leave 2 line spaces
between the figure and the caption. The figure caption is initial caps
and each figure should be numbered consecutively.

Make sure that the figure caption does not get separated from the
figure. Leave extra white space at the bottom of the page rather than
splitting the figure and figure caption.

\begin{figure}[thb]
	\includegraphics[width=.23\textwidth]{figures/mse_by_sample_size_1d.pdf}
	\includegraphics[width=.23\textwidth]{figures/mse_by_sample_size_2d.pdf}
	\includegraphics[width=.23\textwidth]{figures/mse_by_sample_size_3d.pdf}
	\includegraphics[width=.23\textwidth]{figures/mse_by_sample_size_4d.pdf}

	\caption{These are our plots.}
\end{figure}

\subsubsection{Tables}

All tables must be centered, neat, clean, and legible. Do not use hand-drawn tables.
Table number and title always appear above the table.
See Table~\ref{sample-table}.

Use one line space before the table title, one line space after the table title,
and one line space after the table. The table title must be
initial caps and each table numbered consecutively.

\begin{table}[h]
\caption{Sample Table Title} \label{sample-table}
\begin{center}
\begin{tabular}{ll}
\textbf{PART}  &\textbf{DESCRIPTION} \\
\hline \\
Dendrite         &Input terminal \\
Axon             &Output terminal \\
Soma             &Cell body (contains cell nucleus) \\
\end{tabular}
\end{center}
\end{table}

\section{SUPPLEMENTARY MATERIAL}

If you need to include additional appendices during submission, you can include them in the supplementary material file.
You can submit a single file of additional supplementary material which may be either a pdf file (such as proof details) or a zip file for other formats/more files (such as code or videos). 
Note that reviewers are under no obligation to examine your supplementary material. 
If you have only one supplementary pdf file, please upload it as is; otherwise gather everything to the single zip file.

You must use \texttt{aistats2021.sty} as a style file for your supplementary pdf file and follow the same formatting instructions as in the main paper. 
The only difference is that it must be in a \emph{single-column} format.
You can use \texttt{supplement.tex} in our starter pack as a starting point.
Alternatively, you may append the supplementary content to the main paper and split the final PDF into two separate files.

\section{SUBMISSION INSTRUCTIONS}

To submit your paper to AISTATS 2021, please follow these instructions.

\begin{enumerate}
    \item Download \texttt{aistats2021.sty}, \texttt{fancyhdr.sty}, and \texttt{sample\_paper.tex} provided in our starter pack. 
    Please, do not modify the style files as this might result in a formatting violation.

    \item Use \texttt{sample\_paper.tex} as a starting point.
    \item Begin your document with
    \begin{flushleft}
    \texttt{\textbackslash documentclass[twoside]\{article\}}\\
    \texttt{\textbackslash usepackage\{aistats2021\}}
    \end{flushleft}
    The \texttt{twoside} option for the class article allows the
    package \texttt{fancyhdr.sty} to include headings for even and odd
    numbered pages.
    \item When you are ready to submit the manuscript, compile the latex file to obtain the pdf file.
    \item Check that the content of your submission, \emph{excluding} references, is limited to \textbf{8 pages}. The number of pages containing references alone is not limited.
    \item Upload the PDF file along with other supplementary material files to the CMT website.
\end{enumerate}

\subsection{Camera-ready Papers}

%For the camera-ready paper, if you are using \LaTeX, please make sure
%that you follow these instructions.  
% (If you are not using \LaTeX,
%please make sure to achieve the same effect using your chosen
%typesetting package.)

If your papers are accepted, you will need to submit the camera-ready version. Please make sure that you follow these instructions:
\begin{enumerate}
    %\item Download \texttt{fancyhdr.sty} -- the
    %\texttt{aistats2021.sty} file will make use of it.
    \item Change the beginning of your document to
    \begin{flushleft}
    \texttt{\textbackslash documentclass[twoside]\{article\}}\\
    \texttt{\textbackslash usepackage[accepted]\{aistats2021\}}
    \end{flushleft}
    The option \texttt{accepted} for the package
    \texttt{aistats2021.sty} will write a copyright notice at the end of
    the first column of the first page. This option will also print
    headings for the paper.  For the \emph{even} pages, the title of
    the paper will be used as heading and for \emph{odd} pages the
    author names will be used as heading.  If the title of the paper
    is too long or the number of authors is too large, the style will
    print a warning message as heading. If this happens additional
    commands can be used to place as headings shorter versions of the
    title and the author names. This is explained in the next point.
    \item  If you get warning messages as described above, then
    immediately after $\texttt{\textbackslash
    begin\{document\}}$, write
    \begin{flushleft}
    \texttt{\textbackslash runningtitle\{Provide here an alternative
    shorter version of the title of your paper\}}\\
    \texttt{\textbackslash runningauthor\{Provide here the surnames of
    the authors of your paper, all separated by commas\}}
    \end{flushleft}
    Note that the text that appears as argument in \texttt{\textbackslash
      runningtitle} will be printed as a heading in the \emph{even}
    pages. The text that appears as argument in \texttt{\textbackslash
      runningauthor} will be printed as a heading in the \emph{odd}
    pages.  If even the author surnames do not fit, it is acceptable
    to give a subset of author names followed by ``et al.''

    %\item Use the file sample\_paper.tex as an example.

    \item The camera-ready versions of the accepted papers are 8
      pages, plus any additional pages needed for references.

    \item If you need to include additional appendices,
      you can include them in the supplementary
      material file.

    \item Please, do not change the layout given by the above
      instructions and by the style file.

\end{enumerate}

\subsubsection*{Acknowledgements}
All acknowledgments go at the end of the paper, including thanks to reviewers who gave useful comments, to colleagues who contributed to the ideas, and to funding agencies and corporate sponsors that provided financial support. 
To preserve the anonymity, please include acknowledgments \emph{only} in the camera-ready papers.

\subsubsection*{References}

References follow the acknowledgements.  Use an unnumbered third level
heading for the references section.  Please use the same font
size for references as for the body of the paper---remember that
references do not count against your page length total.

\begin{thebibliography}{}
\setlength{\itemindent}{-\leftmargin}
\makeatletter\renewcommand{\@biblabel}[1]{}\makeatother
\bibitem{} J.~Alspector, B.~Gupta, and R.~B.~Allen (1989).
    \newblock Performance of a stochastic learning microchip.
    \newblock In D. S. Touretzky (ed.),
    \textit{Advances in Neural Information Processing Systems 1}, 748--760.
    San Mateo, Calif.: Morgan Kaufmann.

\bibitem{} F.~Rosenblatt (1962).
    \newblock \textit{Principles of Neurodynamics.}
    \newblock Washington, D.C.: Spartan Books.

\bibitem{} G.~Tesauro (1989).
    \newblock Neurogammon wins computer Olympiad.
    \newblock \textit{Neural Computation} \textbf{1}(3):321--323.
\end{thebibliography}

\end{document}

### 2021-02-11 notes/manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
Tampered with directory structure.

--- 2020-10-16 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing.tex
@@
 \documentclass[twoside]{article}
-\usepackage{aistats2021}
 \usepackage{amsmath, amsfonts, amsthm, amssymb}
@@
 \usepackage{verbatim}
+\usepackage{fullpage}
 
@@
 \begin{document}
-
-\twocolumn[
-
-	\aistatstitle{Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs}
-
-	\aistatsauthor{Alden Green, Sivaraman Balakrishnan, and Ryan J. Tibshirani}
-
-	\aistatsaddress{Carnegie Mellon University}]
+\title{Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs}
+\date{\today}
+\maketitle
 
@@
 	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_2d.pdf}
-	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_3d.pdf}\\ 
+	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_3d.pdf} 
 	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_4d.pdf}

### 2021-02-11 manuscripts/laplacian_smoothing/AISTATS/reviews/rebuttal.tex
Another round of re-structuring.

\documentclass{article}

\usepackage{aistats2021_author_response}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % define colors in text
\usepackage{xspace}         % fix spacing around commands

\begin{document}

First of all, we would like to thank all the reviewers for their comments and helpful suggestions. We will begin with a general response that is relevant to multiple reviewers, and then give some specific responses to each reviewer.

\underline{General response} (\textit{``...limited experiment.''}) We agree that the theory would be better complemented by stronger empirical evidence. For a final version, we will extend our experiments to include different regression functions, higher dimensions, the power of our proposed test, and comparisons with other methods for nonparametric regression, especially with the thin-plate spline as suggested by Reviewer 4. We will also improve the presentation of our experimental results, as suggested by Reviewer 6.

\underline{Reviewer 1}
We are glad the reviewer enjoyed the paper.

\underline{Reviewer 2} (\textit{``...how large is d in practice?''}) In practice $d$ is often much larger than 4, while the intrinsic dimension $m$ is usually unknown. However, it is hypothesized (e.g in [Belkin 2006]) that $m$ is often small in real-world problems.

\underline{Reviewer 4}
(\textit{``...intuition behind the difference} [between Laplacian smoothing and thin-plate splines]...'') Laplacian smoothing penalizes the average size of the differences of the function evaluated at the design points, and thus insists on the estimate being on average ``smooth’’ across the design points. Thin-plate splines penalize the average size of the derivative across the domain, and thus are perfectly willing to accept estimates which are very ``wiggly'' near the design points, as long as the wiggliness is sufficiently localized. The key is that for sufficiently wiggly functions, derivatives and differences stop being coupled in high enough dimensions.

\underline{Reviewer 6} (\textit{``...this paper needs to make a more precise point as to its contributions to} [nonparametric regression] \textit{in particular...}) We study the method of Laplacian smoothing, which despite its long history has resisted analysis in the classical statistical setup. We show that although Laplacian smoothing is similar in spirit to first-order smoothing splines when $d = 1$, somewhat surprisingly it avoids the pathological failure of first-order thin-plate splines when $d \geq 2$. Furthermore, we establish that Laplacian smoothing adapts to the manifold hypothesis. Finally, we also show that Laplacian smoothing allows one to make only weak assumptions on “knowing the domain $\mathcal{X}$ and density $p$” while still obtaining minimax rates for Sobolev classes. We will make these points explicitly in our final version.

(\textit{“...a more precise comparison with other methods...”}) While it is possible to derive other minimax optimal regressors in our setup---e.g. by tuning kernel smoothing---our goal is somewhat different: to consider Laplacian smoothing, a widely used method with a strong empirical track record, and show that it specifically has optimal statistical properties. However, we agree with the reviewer that a more detailed comparison could be useful and will include this in the final version. Our analysis differs substantially from standard analyses of classical non-parametric methods (kernel smoothing, RKHS regression and truncated series estimators), and we note some important distinctions. \textit{Kernel smoothing}: the bias and variance terms we derive for Laplacian smoothing are quite different from those typical to kernel smoothing (even if the worst-case rates match), and indicate conditions under which Laplacian smoothing will outperform kernel smoothing. \textit{RKHS regression}: analyses of variational methods similar in spirit to Laplacian smoothing---i.e. thin-plate splines, but also more generally kernel ridge regression---typically assume that $f_0$ lies close to an RKHS; when $d \geq 2$ the first-order Sobolev space $H^{1}(\mathcal{X})$ is not an RKHS, and most standard analyses do not apply without careful modifications. \textit{Truncated series estimators}: the analyses of least squares over a subspace of $L^2(\mathcal{X})$ typically rely on an \textit{a priori} bound on the approximation error---which is hard to obtain in the manifold case without “knowing” the manifold in a strong sense.  In contrast, for Laplacian smoothing we derive a data-dependent bound on the approximation error, and upper bound this by a population-level quantity with high probability.

(\textit{“...practical implementation of the threshold.”}) In the final paper, we will make clear that the level-$\alpha$ permutation threshold is computed as follows: (1) randomly draw $M$ permutations $\pi_1,...,\pi_M$ of the indices $\{1,...,n\}$; (2) for each $m = 1,...,M$, compute the test statistic $T_m$ with respect to the permuted responses $(Y_{\pi_m(1)},X_1),...,(Y_{\pi_m(n)},X_n)$; (3) set the threshold to be the $(1 - \alpha)$-quantile of the empirical distribution of $\{T_1,...,T_M\}$.

(\textit{“...adaptivity to the intrinsic dimension m...”}) In our Theorems 3 and 4 the choice of radius $r$ depends on $m$. Typically, $m$ is unknown: although it can be estimated, practically we would recommend to choose $r$ by cross-validation. We believe, arguing along the lines of [Gyorfi 2006], that such a choice will achieve the same rates adaptively.

(\textit{“... intuition as to how your assumptions are used in your proofs?”}) In our final paper, we will include the following comments on assumptions. The Lipschitz assumptions, on the density and kernel, allow us to relate nonlocal approximations of the Sobolev semi-norm to the Sobolev semi-norm. The assumption that $K(c) > 0$ for some fixed $c > 0$ is fundamental---otherwise the radius $r$ does not truly reflect the scale of connectivity; on the other hand the requirement that the kernel be compactly supported on $[0,1]$ is for technical convenience. The lower bound on the density is used for comparing sample- and population-level eigenvalues; we make use of upper bounds on the optimal-transport distance in infinity norm between $P_n$ and $P$,  and such distance can be very large when the tails of $p(x)$ are close to $0$. Finally, the lower and upper bounds on the density are required for the lower and upper bounds in (16)---Weyl’s Law---to hold.

\end{document}

### 2021-02-11 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Another round of re-structuring.

--- 2020-09-04 notes/graph_regression2.tex
@@
 \begin{document}
-\title{Minimax-optimal nonparametric regression with graph Laplacians}
+\title{Minimax-optimal regression over Sobolev Spaces with Neighborhood Graphs}
 \author{Alden Green}
@@
 
-\section{Introduction}
-\label{sec:introduction}
-
-In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ with support $\Xset \subset \Rd$, and
-\begin{equation}
-\label{eqn:random_design_regression}
-Y_i = f_0(X_i) + \varepsilon_i
-\end{equation}
-with independent Gaussian noise $\varepsilon_i \sim N(0,1)$.
-Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether or not $f_0 = 0$, i.e whether or not there is a signal present.
-
-In graph based nonparametric regression, the idea is to perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$.
-For a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$.
-The weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ is an $n \times n$ matrix with entries
-\begin{equation*}
-\label{eqn:neighborhood_graph}
-{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr),
-\end{equation*}
-and the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$.
-Our star operator, the neighborhood graph Laplacian, can then be written as
-\begin{equation}
-\label{eqn:graph_Laplacian}
-\Lap_{n,r} = \bf{D} - \bf{W}
-\end{equation}
-The two graph-based regression methods we will consider are \emph{Laplacian eigenmaps} and \emph{Laplacian smoothing}.
-Each method uses $\Lap_{n,r}$ in a different manner to construct an estimator of $f_0$.
-Letting $v_1(G_{n,r}),\ldots,v_n(G_{n,r})$ denote eigenvectors of $\Lap_{n,r}$, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ \citep{belkin2003} is given by
-\begin{equation}
-\label{eqn:laplacian_eigenmaps}
-\wh{f}_{\textrm{LE}} = \sum_{k = 1}^{\kappa} \Dotp{Y}{v_k(G_{n,r})}_n v_{k}(G_{n,r})
-\end{equation}
-where $\kappa \in \{1,...,n\}$ is a tuning parameter.
-(Here $\dotp{u}{v}_n = \frac{1}{n}\sum_{i = 1}^{n} u_i v_i$ for vectors $u,v \in \Reals^n$.) The Laplacian smoothing estimator $\wt{f}_{\LS}$ \citep{smola2003} is a penalized least square estimator, given by
-\begin{equation}
-\label{eqn:laplacian_smoothing}
-\wt{f}_{\LS}= \min_{f \in \Reals^n} \Bigl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_{n,r}^s f \Bigr\}
-\end{equation}
-where $\rho > 0$ acts a tuning parameter $f^T \Lap_{{n,r}}^s f$.
-Then, assuming~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are reasonable estimators of $f_0$, the squared empirical norms
-\begin{equation}
-T_{\LE} = \Bigl\|\wh{f}_{\LE}\Bigr\|_n^2 \label{eqn:laplacian_eigenmaps_test}
-\end{equation}
-and
-\begin{equation}
-T_{\LS} = \Bigl\|\wt{f}_{\LS}\Bigr\|_n^2 \label{eqn:laplacian_smoothing_test}
-\end{equation}
-are in turn reasonable statistics to test whether or not $f_0 = 0$.
-
-\begin{itemize}
-	\item \textbf{Motivate graph Laplacians.} There already exist a vast number of methods for both the nonparametric testing and, especially, the nonparametric estimation problems.
-In fact, as we will see in Section~\ref{sec:problem_setup_and_background} some of these methods are actually quite analogous to the graph Laplacian estimation and testing methods we have just introduced.
-Nevertheless, methods such as~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} have a deserved place in the nonparametric regression toolbox.
-Their advantages include:
-	\begin{itemize}
-		\item Simplicity.
-		\item Computational efficiency and stability.
-		\item Domain and density adaptivity.
-		\item Easily generalizable.
-		\item Practical success.
-	\end{itemize}
-	\item \textbf{Our contributions, short version.} For these reasons, there is a  substantial body of work assessing the statistical properties of graph Laplacian based methods.
-\textcolor{red}{(TODO): Include citations.} In this paper, we adopt a different perspective, and analyze~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} through the minimax lens.
-We show that when the regression function $f_0$ is suitably smooth (by which we mean that it belongs to a first-order Sobolev ball; roughly speaking, that it possesses weak derivative $Df$ bounded in $L^2(\Xset)$ norm) both~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are minimax optimal estimators, and both~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} are minimax optimal test statistics.
-We also show that under appropriate conditions, these estimators and test statistics can leverage additional assumed structure, either on the regression function $f_0$ or distribution $P$.
-	\item \textbf{Organization.}
-\end{itemize}
-
-\paragraph{Notation.}
-\begin{itemize}
-	\item We write $\mathbb{Z}^d$ for $\mathbb{Z} \otimes \cdots \mathbb{Z}$, where the tensor product is taken a total of $d$ times.
-	\item We denote by $\Leb^2(\Xset)$ be the set of square integrable functions $f:\Xset \to \Reals$.
-	\item For functions $f,g$ which map from $\Xset$ to $\Reals$, we write $\dotp{f}{g}_n := \frac{1}{n}\sum_{i = 1}^{n} f(X_i) g(X_i)$, which we refer to as the \emph{empirical inner product}, and $\norm{f}_n^2 = \dotp{f}{f}_n$, which we refer to as the (squared) \emph{empirical norm}.
-For vectors $v \in \Reals^n$, we write $\norm{v}_p^p := \sum_{i = 1}^{n} \abs{v_i}^p$.
-	\item For a graph $G$, we use $\Lap_G$ for the Laplacian of $G$.
-The matrix $\Lap_G$ is a square, symmetric, positive semi-definite matrix, with $n$ eigenvalues $\lambda_k(G)$ and eigenvector $v_k(G)$ pairs, defined according to
-	\begin{equation*}
-	\Lap_G v_k(G) = \lambda_k(G) v_k(G), ~~\frac{1}{n} \norm{v_k(G)}_2^2 = 1 
-	\end{equation*}
-	and ordered so that $0 \leq \lambda_1(G) \leq \ldots \leq \lambda_n(G)$. For convenience, we write $\Lap_{n,r}$ instead of $\Lap_{G_{n,r}}$. 
-\end{itemize}
-
-\section{Problem Setup and Background}
-\label{sec:problem_setup_and_background}
-
-Before we get to our main results, we will review some of the relevant literature on minimax optimal nonparametric estimation and testing of Sobolev functions.
-We will also briefly recall some (of the many) estimators and tests which achieve these rates: in particular the \emph{truncated series} and \emph{smoothing spline} methods.
-As we will see, Laplacian eigenmaps and Laplacian smoothing can be seen as graph-based, i.e.
-discrete, analogues to these ``continuuum'' methods.
-
-\subsection{Sobolev spaces}
-\label{subsec:sobolev_spaces}
-
-\begin{itemize}
-	\item \textbf{Sobolev spaces.}  Roughly speaking, Sobolev spaces contain functions $f \in \Leb^p(\Xset)$ with derivatives which themselves belong to $\Leb^p(\Xset)$.
-More formally, for given integers $s, p > 0$, the Sobolev space $W^{s,p}(\Xset)$ consists of all functions $f \in \Leb^p(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^p(\Xset)$.
-The Sobolev-$\{s,p\}$ seminorm $\seminorm{f}_{W^{s,p}(\Xset)}$ and norm $\norm{f}_{W^{s,p(\Xset)}}$ are then given by
-	\begin{equation*}
-	\seminorm{f}_{W^{s,p}(\Xset)}^p = \sum_{\abs{\alpha} = s}\int_{\mathcal{X}} \Bigl|\bigl(D^{\alpha}f\bigr)(x)\Bigr|^p \,dx, ~~ \norm{f}_{W^{s,p}(\Xset)}^p = \sum_{k = 0}^{s} \seminorm{f}_{W^{k,p}(\Xset)}^p
-	\end{equation*}
-	and for a given $M > 0$, the corresponding ball is $W^{s,p}(\Xset, M) = \set{f: \norm{f}_{W^{s,p}(\Xset)} \leq M}$.
-In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space, and we adopt the usual convention of writing $H^s(\Xset) = W^{s,p}(\Xset)$ and $H^s(\Xset,M) = W^{s,2}(\Xset,M)$.
-We will confine our attention to this special case hereafter.
-	\item \textbf{Connections to sequence spaces.} The balls $H^s(\Xset,M)$ are closely connected with the \emph{Sobolev ellipsoids} $\Theta_s(M) \subseteq \ell^2(\mathbb{Z}^d)$, defined as
-	\begin{equation*}
-	\Theta_s(M) = \Bigl\{\theta \in \ell_2(\mathbb{Z}^d): \sum_{k \in \mathbb{Z}^d} \theta_k^2 \abs{k}^{2s} < M\Bigr\}
-	\end{equation*}
-	Let us assume that the domain $\Xset = [0,1]^d$. Letting be $\set{\phi_k}_{k \in \mathbb{Z}^d}$ be the tensor product trigonometric polynomial basis of $\Leb^2(\Xset)$, the space
-	\begin{equation*}
-	H_{\textrm{per}}^s(\Xset,M) = \Bigl\{\sum_{k \in \mathbb{Z}^d} \theta_k \phi_k: \theta_k \in \Theta_s(M) \Bigr\}
-	\end{equation*}
-	satisfies $H_{\textrm{per}}^s(\Xset,M) \subseteq H^s(\Xset,M')$ for a sufficiently large value of $M_1 > 0$.
-On the other hand, any $1$-periodic function $f \in W^{s,p}(\Xset,M)$ also belongs to $H_{\textrm{per}}^s(\Xset,M_0)$ for a sufficiently large value of $M_0 > 0$.
-	\item \textbf{Zero-trace Sobolev spaces.} We will also be interested in the \emph{zero-trace} Sobolev spaces.
-\end{itemize}
-
-\subsection{Orthogonal series and smoothing splines.}
-\label{subsec:continuum_methods}
-When the regression function is assumed to belong to a Sobolev space, two approaches particularly well-suited for nonparametric regression are orthogonal series \textcolor{red}{projection}, and smoothing splines.
-The orthogonal series estimator $\wh{f}_{\OS}$ is given by
-\begin{equation}
-\label{eqn:orthogonal_series}
-\wh{f}_{\OS} = \sum_{k: \abs{k} \leq \kappa} \dotp{Y}{\phi_k}_n \phi_k .
-\end{equation}
-where $\kappa \geq 0$ is a hyperparameter, and we recall that $\{\phi_k\}_{k \in \mathbb{Z}^d}$ is the tensor product trigonometric polynomial basis of $\Leb^2(\Xset)$.
-The smoothing spline estimator $\wt{f}_{\SM}$ is a variational estimator, defined as the solution to the following optimization problem,
-\begin{equation}
-\label{eqn:smoothing_spline}
-\wt{f}_{\SM} = \argmin_{f} \sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \seminorm{f}_{H^{s}(\Xset)}^2
-\end{equation}
-where the minimum is taken over all $f \in H^s(\Xset)$.
-
-\begin{itemize}
-	\item \textbf{Connecting orthogonal series and Laplacian eigenmaps estimators.}The orthogonal series estimator $\wh{f}_{\OS}$ is closely connected with the Laplacian eigenmaps estimator $\wh{f}_{\LE}$.
-At a high-level, one can go from~\eqref{eqn:orthogonal_series} to~\eqref{eqn:laplacian_eigenmaps} simply by substituting $\{v_k\}_{k \in [n]}$ for $\{\phi_k\}_{k \in \mathbb{Z}^d}$.
-More deeply, just as the vectors $v_k$ are eigenvectors of the graph Laplacian $\Lap_{n,r}$, so too are the functions $\phi_k$ eigenfunctions of a particular differential operator $\Delta: C^2(\Xset) \to \Leb^2(\Xset)$, known as the \emph{Laplace} operator and formally defined as
-	\begin{equation}
-	\label{eqn:laplace_operator}
-	\Delta g := -\mathrm{div}(\nabla g).
-	\end{equation}
-	In fact, it has been shown in \textcolor{red}{(references)} that when the radius $r \to 0$ as $n \to \infty$, the graph Laplacian $\Lap_{n,r}$ converges in an appropriate sense to a weighted version $\Delta_{P}$ of $\Delta$, with
-	\begin{equation}
-	\Delta_{P} g := -\frac{1}{p}\mathrm{div}(p^2 \nabla g).
-	\end{equation}
-	Along similar lines, \textcolor{red}{(references)} have established that for any fixed $k \geq 0$, as $n \to \infty$ the eigenvector $v_k$ will converge, again in an appropriate sense, to an eigenfunction $\phi_k$ of $\Delta_{P}$.
-The upshot is that when $P$ is the uniform measure, for an appropriate sequence of radii $r(n) \to 0$ and any fixed $\kappa \geq 0$ the estimator $\wh{f}_{OS}$ is in fact the limit of $\wh{f}_{\LE}$, in the sense that
-	\begin{equation}
-	\label{eqn:laplacian_eigenmaps_consistency}
-	\abs{\wh{f}_{\LE}(X_i) - \wh{f}_{\OS}(X_i)} \overset{P}{\to} 0~~\textrm{as $n \to \infty$.}
-	\end{equation} 
-	for each $i \in \mathbb{N}$.
-	\item 
-	\textbf{Connecting smoothing spline and Laplacian smoothing estimators.} In a similar vein, the smoothing spline estimator $\wh{f}_{\SM}$ has deep connections with the Laplacian smoothing estimator $\wh{f}_{\LS}$.
-Using the definition~\eqref{eqn:laplace_operator}, recursively applying integration by parts, and then invoking the aforementioned convergence properties of $\Lap_{n,r}$, it can be shown that \textcolor{red}{references} that for any $f \in C^{2s}(\Xset)$,
-	\begin{equation}
-	\label{eqn:pointwise_seminorm_convergence}
-	\abs{f}_{H^s(\Xset)} = \int_{\Xset} \bigl(\Delta^sf\bigr)(x) f(x) \,dx = \lim_{n \to \infty} f^T \Lap_{n,r}^s f.
-	\end{equation}
-	By applying appropriation concentration inequalities, one can extend~\eqref{eqn:pointwise_seminorm_convergence} to hold uniformly over all $f \in H^{2s}(\Xset)$.
-While this does not imply a formal statement about the convergence of $\wh{f}_{\LS}$ to $\wh{f}_{SM}$ as in~\eqref{eqn:laplacian_eigenmaps_consistency}---note the difference between the domain of minimization in~\eqref{eqn:smoothing_spline} and the regularity required for~\eqref{eqn:pointwise_seminorm_convergence}---it nevertheless makes clear that the estimators $\wh{f}_{\LS}$ and $\wh{f}_{\SM}$ are linked.
-	\item \textbf{Domain and density adaptivity.} \textcolor{red}{(TODO)}
-\end{itemize}
-
-\subsection{Minimax rates over Sobolev classes}
-\label{subsec:minimax_rates}
-One reason all of this is intriguing is that, as we mentioned previously, the continuum estimators $\wh{f}_{\OS}$ and $\wh{f}_{\SM}$---and test statistics based on functionals of these estimators--- have strong statistical properties.
-In particular, these estimators and tests are known to have error rates within a constant factor of (minimax) optimal.
-
-\paragraph{Estimation rates.}
-Under appropriate regularity conditions, the minimax estimation rate over Sobolev balls is 
-\begin{equation}
-\label{eqn:sobolev_space_estimation_minimax_rate}
-\inf_{\wh{f}} \sup_{f_0 \in \mc{H}^s(M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d \geq 0$,}
-\end{equation}
-where the infimum is over all measurable maps $\wh{f}(X_1,Y_1,\ldots,X_n,Y_n)$, and the function class $\mc{H}^s(M)$ is one of $H^s(\Xset,M)$, $H_{\mathrm{per}}^s(\Xset,M)$ or $H_0^s(\Xset,M)$.
-Orthogonal series and smoothing spline estimators are two (of the many) estimators which certify the upper bound in~\eqref{eqn:sobolev_space_estimation_minimax_rate}.
-Concretely, if we are willing to assume
-\begin{enumerate}[label=(P\arabic*)]
-	\item
-	\label{asmp:uniform_density}
-	the distribution $P$ is the uniform probability measure defined on domain $\Xset = [0,1]^d$.
-\end{enumerate}
-then the orthogonal series estimator is minimax optimal over the Sobolev ellipsoids $H_{\mathrm{per}}^s(M)$, meaning
-\begin{equation*}
-\sup_{f \in H_{\mathrm{per}}^s(\Xset,M)}\Ebb\Bigl[\norm{\wh{f}_{\OS} - f}_{L^2(\Xset)}^2\Bigr]  \lesssim n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d \geq 0$.}
-\end{equation*}
-An advantage of $\wt{f}_{\SM}$ over $\wh{f}_{\OS}$ is that it requires less stringent assumptions on the design density in order to be minimax optimal.
-Instead of Assumption~\ref{asmp:uniform_density}, we will merely assume that the density is bounded above and below.
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{1}
-	\item 
-	\label{asmp:bounded_density}
-	The distribution $P$ is defined on domain $\Xset = [0,1]^d$, and admits a density $p$ with respect to the uniform (Lebesgue) measure on $[0,1]^d$.
-The density $p$ is bounded away from $0$ and $\infty$, that is
-	\begin{equation*}
-	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty
-	\end{equation*}
-	for all $x \in [0,1]^d$.
-\end{enumerate}
-The assumption~\ref{asmp:bounded_density} is itself a standard but significant assumption; we will return to discuss it further at \textcolor{red}{a later point}.
-Assuming~\ref{asmp:bounded_density}, the estimator $\wt{f}_{\SM}$ satisfies
-\begin{equation*}
-\sup_{f_0 \in H^s(\Xset,M)}\Ebb\Bigl[\norm{\wh{f}_{\SM} - f_0}_{L^2(\Xset)}^2\Bigr]  \lesssim n^{-2s/(2s + d)}~~\textrm{for all $2s > d$.}
-\end{equation*}
-The restriction $2s > d$ is a severe limitation of smoothing splines relative to  orthogonal series.
-It is worth emphasizing that this is a fundamental defect of the smoothing spline estimator, as opposed to any flaw in the theory; when $2s \leq d$, the optimization problem~\eqref{eqn:smoothing_spline} is not even well-defined!
-\textcolor{red}{(TODO):} succinctly explain why.
-
-\paragraph{Goodness-of-fit testing rates.}
-
-\begin{itemize}
-	\item \textbf{Minimax hypothesis testing.} In the nonparametric regression hypothesis testing problem, we ask for a test function (formally, a Borel measurable map $\phi: \mc{X}^n \times \Reals^n \to \{0,1\}$) which can distinguish between the null and alternative hypotheses
-	\begin{equation}
-	\mathbf{H}_0: f_0 = 0, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \neq 0.
-	\end{equation}
-	We measure the performance of a test $\phi$ by its Type I error $\Ebb_0(1 - \phi)$, and Type II error $\mathbb{E}_{f_0}(\phi)$.
-It is well known that no test $\phi$ with controlled Type I error---say, $\Ebb_0(\phi) \leq 1/b$ for some constant $b > 1$---
-	can have controlled Type II error---$\Ebb_{f_0}(\phi) \leq 1/b$--- for every $f_0 \neq 0$.
-As a result, it is standard to impose two kinds of restrictions on the alternative hypothesis class $\mathbf{H}_a$.
-The first is a smoothness condition on $f_0$: specifically, we will assume that $f_0$ belongs to a Sobolev ball $\mc{H}^s(M)$.
-The latter is a separation condition, enforcing a minimum distance between $f_0$ and $0$: specifically, we will insist that $\norm{f_0  - 0}_{\Leb^2(\Xset)} > \epsilon$ for some $\epsilon > 0$.
-	\item\textbf{Minimax testing rates over Sobolev classes.} Once smoothness and separation conditions are established, the following becomes a natural question to ask: what is the smallest value of $\epsilon$ for which there exists a test $\phi$ with both small Type I error, and small Type II error over all functions $f_0$ which are smooth and separated from $0$ by at least $\epsilon$?
-This smallest separation $\epsilon$ is known as the critical radius, and in the case of the Sobolev classes it is given by
-	\begin{equation*}
-	\epsilon_n^{\star}\bigl(\mc{H}^s(M)\bigr):= \inf\biggl\{\epsilon: \inf_{\phi} \Bigl\{\Ebb_0(\phi) +  \sup_{\substack{f_0 \in \mc{H}^s(M) \\ \norm{f_0}_{\Leb^2(\Xset)} \geq \epsilon}} \Ebb_{f_0}[1 - \phi]\Bigr\} \leq \frac{2}{b} \biggr\} \asymp n^{-2s/(4s + d)}
-	\end{equation*}
-	for any $4s > d$, and $\mc{H}^s(M)$ one of $H^s(\Xset,M)$, $H_{\mathrm{per}}^s(\Xset,M)$ or $H_0^s(\Xset,M)$. On the other hand, when $4s \leq d$ the minimax rate is dimension-free,
-	\begin{equation*}
-	\epsilon_n^{\star}\Bigl(\mc{H}^s(M)\Bigr) \asymp n^{-1/4},
-	\end{equation*}
-	revealing a distinction between the estimation and goodness-of-fit testing problems.
-	\item \textbf{Error rates of orthogonal series and smoothing spline-based tests.} 
-	Just as $\wh{f}_{\OS}$ and $\wh{f}_{\SM}$ achieve minimax optimal estimation rates, tests using the plug-in statistics
-	\begin{equation*}
-	T_{\OS} = \bigl\|\wh{f}_{\OS}\bigr\|_{\Leb^2(\Xset)}^2,~~\textrm{and}~~T_{\SM} = \bigl\|\wt{f}_{\SM}\bigr\|_{n}^2
-	\end{equation*}
-	are minimax optimal for the goodness-of-fit testing problem over Sobolev balls (see \citet{ingster2009} for analysis of the former statistic, and \citet{liu2019} for analysis of the latter).
-\end{itemize}
-
-\subsection{Our contributions}
-\label{subsec:our_contributions}
-
-\begin{itemize}
-	\item \textbf{Departure from what's already known.} At a high level, the connections between the graph-based methods~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} and the continuum methods outlined in Subsection~\ref{subsec:continuum_methods} give reason to believe that the strong statistical properties of the latter are shared by the former.
-Making this rough statement precise is the content of this paper.
-In particular, we will prove the following results:
-	\begin{itemize}
-		\item \textbf{Minimax optimal estimation.}
-		We show that with high probability, $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2s/(2s + d)}$ over all $f_0 \in H^1(\Xset;M)$.
-The same conclusions hold with respect to $\wt{f}_{\LS}$ when $d \leq 4$.
-		\item \textbf{Minimax optimal testing.}
-		We show that with high probability, a test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in H^1(\Xset;M)$ satisfies $\norm{f_0}_{\Leb^2(\Xset)} \gtrsim \max\{n^{-2s/(4s + d)},n^{-1/4}\}$.
-\textcolor{red}{The same conclusions hold with respect to $\wt{f}_{LS}$.}
-		\item \textbf{Taking advantage of higher-order smoothness conditions.} Under appropriate boundary conditions on the function classes $H^s(\Xset)$ and additional smoothness conditions on the density $p$, when $s > 1$  each of the aforementioned conclusions hold for certain values of $d$.
-		\item \textbf{Manifold adaptivity.}
-		If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$, each of the aforementioned rates hold with $d$ replaced by $m$.
-	\end{itemize}	
-\end{itemize}
-
-\section{Graph Sobolev classes}
-\label{sec:graph_sobolev_classes}
-
-\begin{itemize}
-	\item For $s,M > 0$, define the order-$s$ graph Sobolev ball to be
-	\begin{equation}
-	\label{eqn:graph_sobolev_ball}
-	H^s(G_{n,r},M) = \set{f \in \Reals^n: f^T \Lap_{n,r}^s f \leq M}
-	\end{equation}
-	where we recall that the \emph{graph Sobolev seminorm} $f^T \Lap_{{n,r}}^s f$ is precisely the penalty term in~\eqref{eqn:laplacian_smoothing}. 
-	To understand the relevance of the graph Sobolev ball to the analysis of our graph-based procedures, let us reason again by analogy with the continuum world.
-The statistical analysis of continuum-time methods such as Fourier series projection and smoothing splines rests on two facts about the continuum classes $\mc{H}^s(\Xset)$: first, the \textit{a priori} assumption that the regression function $f_0$ belongs to an appropriate Sobolev ball $\mc{H}(\Xset,M)$, and second that the asymptotic decay of the eigenvalues of the continuum Laplace operator $\Delta$ obey Weyl's law, meaning $\lambda_k(\Delta) \asymp \abs{k}^{2s}$ for $k \in \mathbb{Z}^d$.
-In this section, we will show that both these statements have appropriate analogues in the graph setting.
-Roughly speaking, we will show first that for functions $f$ belonging to $\mc{H}^s(\Xset, M)$, the evaluations $f = (f(X_1),\ldots,f(X_n)) \in \Reals^n$ satisfy the upper bound
-	\begin{equation}
-	\label{eqn:graph_sobolev_seminorm_asymptotic}
-	f^T \Lap_{n,r}^s f \lesssim M(n^{s + 1}r^{s(d + 2)})
-	\end{equation}
-	and second that the eigenvalues $\lambda_k(G_{n,r})$ satisfy the lower bound
-	\begin{equation*}
-	\bigl[\lambda_k(G_{n,r})\bigr]^s = \lambda_k(\Lap_{{n,r}}^s) \gtrsim k^{2s/d} (n^s r^{s(d + 2)});
-	\end{equation*}
-	both statements will hold with high probability.
-These results elucidate why our graph-based procedures ``work'' as well as their continuum counterparts---in the sense of being statistically rate-optimal over (continuum) Sobolev classes.
-They also clarify when our analysis does and does not apply---in terms of various degrees of smoothness $s$ and dimensions $d$.
-\end{itemize}
-
-\subsection{Graph Sobolev seminorm}
-We will state and prove three different results involving the graph Sobolev semi-norm $f_0 \Lap_{n,r}^s f_0$.
-In each case, we will derive that the asymptotic relation~\eqref{eqn:graph_sobolev_seminorm_asymptotic} holds with high probability, whenever $r \to 0$ sufficiently slowly as $n \to \infty$.
-The difference between the three results will be the assumptions we make, and the resulting rate of permissible decay for the radius $r$.
-
-\subsubsection{First-order graph Sobolev semi-norm.}
-We begin by upper bounding $f_0^T \Lap_{n,r} f_0$ under the assumption $f_0 \in H^1(\Xset,M)$.
-This will be the easiest of the three Lemmas to prove, while still giving some sense of how the proofs develop in the more complicated cases.
-The upper bound will depend on the kernel $K$; we assume throughout that $K:[0,\infty) \to [0,\infty)$ is a $K_{\max}$-Lipschitz function supported on $[0,1]$, and without loss of generality that
-\begin{equation*}
-\int_{\Reals^d} K(\norm{z}) \,dz = 1.
-\end{equation*}
-\begin{lemma}
-	\label{lem:first_order_graph_sobolev_seminorm}
-	Suppose $\Xset$ is a Lipschitz domain, and that $p(x) \leq p_{\max}$ for all $x \in \Xset$.
-Then, for any $f \in H^1(\Xset,M)$, there exists a constant $C > 0$ which may depend only on $\Xset$ and $d$ such that
-	\begin{equation}
-	f^T \Lap_{n,r} f \leq \frac{C}{\delta} n^2 r^{d + 2} M^2
-	\end{equation}
-	with probability at least $1 - \delta$.
-\end{lemma}
-
-\subsubsection{General graph Sobolev semi-norm}
-
-When $s > 1$, in order to obtain the desired upper bound on the graph Sobolev semi-norm we must make some additional assumptions regarding the behavior of $f_0$ near the boundary of $\Xset$, as well as on the smoothness of the density function $p$.
-\begin{theorem}
-	\label{thm:graph_sobolev_seminorm}
-	For some integer $s > 1$, suppose that the density $p \in C^{s - 1}(\Xset,p_{\max})$ and $f \in H_0^s(\Xset,M)$, and additionally that the domain $\Xset$ is Lipschitz.
-Let $r = r(n)$ be a sequence of connectivity radii satisfying $r(n) \to 0$ and $r(n) \gtrsim n^{-1/(2(s - 1) +d)}$ as $n \to \infty$.
-Then for all $n$ sufficiently large, there exists a constant $C > 0$ which may depend only on $\Xset$ and $d$ such that
-	\begin{equation}
-	\label{eqn:graph_sobolev_seminorm}
-	f^T \Lap_{n,r}^s f \leq \frac{C}{\delta} n^{s + 1} r^{s(d + 2)}M^2
-	\end{equation}
-	with probability at least $1 - \delta$.
-\end{theorem}
-It is worth pausing to make a few remarks.
-\begin{itemize}
-	\item \textbf{Bounded norm vs. bounded semi-norm.}
-	\item \textbf{Markov's inequality, as opposed to exponential concentration.}
-	\item \textbf{Simpler proof and stronger concentration under Holder assumption.}
-	\item \textbf{Explain why~\eqref{eqn:graph_sobolev_seminorm} should be familiar, and a lower bound on $r$ expected.} The scaling (in $n$, $r$, and $M$) on the right hand side of~\eqref{eqn:graph_sobolev_seminorm} will be familiar to the informed reader.
-For a sufficiently regular function $f \in C^3(\Xset,M)$ and density $p \in C^1(\Xset,p_{\max})$, it is known \textcolor{red}{(references)} that $\frac{1}{nr^{d + 2}}\Lap_{n,r}f \to \Delta_pf$ for \textcolor{red}{$r \geq ???$}.
-\textcolor{red}{A recursive argument shows that when $f \in C^{2s + 1}(\Xset)$ and $p \in C^{2s}(\Xset)$ are sufficiently regular and $r$ is sufficiently large},
-	\begin{equation*}
-	\Bigl(\frac{1}{nr^{d + 2}}\Lap_{n,r}\Bigr)^s f \to \Delta_p^sf \Longrightarrow \frac{1}{n^{s + 1} r^{s(d + 2)}} f^T \Lap_{n,r}^s f \to \int_{\Xset} f(x) \cdot  \Delta_p^sf(x) p(x) \,dx \leq c M^2.
-	\end{equation*}
-	\item \textbf{Explain why the actual proof is more subtle.} In fact, the actual proof of Theorem~\ref{thm:graph_sobolev_seminorm} is rather more involved.
-Among other subtleties we highlight the mismatch, in the previous remark, between the number of smooth derivatives $f$ possesses ($2s + 1$) and the degree of the Laplacian operator $\Lap_{{n,r}}^s$ $(s)$; this is unsuitable for our purposes.
-A more careful analysis implies~\eqref{thm:graph_sobolev_seminorm}, but note well that~\eqref{eqn:graph_sobolev_seminorm} is only an upper bound rather than a statement about consistency.
-This is the price we pay for insisting that the assumed number of smooth derivatives match the degree of the Laplacian; fortunately an upper bound will be sufficient for our purposes.
-\end{itemize}
-
-\subsubsection{Graph Sobolev semi-norm under manifold assumption.}
-
-\textcolor{red}{(TODO)}
-
-\subsection{Graph Laplacian eigenvalues}
-
-\textcolor{red}{(TODO)}
-
-\section{Minimax optimal graph Laplacian methods}
-\label{sec:minimax_optimal_graph_Laplacian_methods}
-
-Let us start by assuming $f_0 \in H^1(P,M)$.
-In this section, we will establish that the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and the resulting test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test}, achieve the standard minimax optimal estimation and hypothesis testing rates, respectively.
-Pleasingly, these results will go through with essentially no additional regularity conditions---additional, that is, to~\ref{asmp:bounded_density}---on either the regression function $f_0$ or the density $p$.
-The story will also be complete: we will obtain the ``right'' rates for all regimes--that is, all values of $s$ and $d$---for which we would expect to get these rates.
-
-\subsection{Minimax optimal estimation}
-
-Under appropriate conditions, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves minimax rates over all $f_0 \in H^1(P,M)$ with high probability.
-The conditions we impose will be on the neighborhood graph $G_{n,r}$.
-As the construction of the graph is in the user's control, these assumptions are therefore mild.
-\begin{enumerate}[label=(K\arabic*)]
-	\item 
-	\label{asmp:kernel_form}
-	The kernel function $K(z) = \1\Bigl\{\norm{z}_2/r \leq 1 \Bigr\}$.
-	\item 
-	\label{asmp:kernel_radius_estimation}
-	The radii $r = r(n)$ satisfies $r(n) \to 0$ as $n \to \infty$, and additionally
-	\begin{equation*}
-	n^{-1/d}\leq r(n) \leq M^{2/(2 + d)}n^{-1/(2 + d)}
-	\end{equation*}
-	for all $n$ sufficiently large.
-\end{enumerate}
-
-We use~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_estimation} to ensure that the approximation error (bias) of $\wh{f}_{\LE}$ is of the correct order.
-On the other hand, bounding the estimation error of $\wh{f}_{\LE}$ -- both in expectation and in probability-- is straightforward.
-Together, bounds on these two sources of error imply Theorem~\ref{thm:laplacian_eigenmaps_estimation1}.
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_estimation1}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
-Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_estimation}.
-Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
-	\begin{equation*}
-	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4 + d)} n^{-2/(2 + d)}
-	\end{equation*}
-	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
-\end{theorem}
-
-\begin{itemize}
-	\item Discuss the choice of radius $r$ imposed by~\ref{asmp:kernel_radius_estimation}.
-	\item Comment on the distinction between error rates as measured by the norm in $L^2(P_n)$ vs. $L^2(P)$.
-	\item \textbf{Proof techniques}: to prove results we rely on both (a) classical bias-variance decomposition of orthogonal series estimators (see e.g.
-\textcolor{red}{Tsybakov}), and (b) pointwise and variational results on neighborhood graph Laplacian semi-norms and eigenvalues (see e.g.
-\textcolor{red}{Belkin04, Belkin08, GarciaTrillos18, Calder19})
-\end{itemize}
-
-Theorem~\ref{thm:laplacian_smoothing_estimation1} is analogous to Theorem~\ref{thm:laplacian_eigenmaps_estimation1}, except it deals with the Laplacian smoothing estimator $\wt{f}_{\LS}$ rather than the Laplacian eigenmaps estimator $\wh{f}_{\LE}$.
-\begin{theorem}
-	\label{thm:laplacian_smoothing_estimation1}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
-Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_estimation}.
-Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when we set $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$ the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
-	\begin{equation*}
-	\Bigl\|\wh{f}_{\LS} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4 + d)} n^{-2/(2 + d)}
-	\end{equation*}
-	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
-\end{theorem}
-
-\begin{itemize}
-	\item \textbf{Proof techniques}: Because the solution to~\eqref{eqn:laplacian_smoothing} has an analytic form, we can rely on a bias-variance decomposition---as opposed to empirical process theory---to derive~Theorem~\ref{thm:laplacian_smoothing_estimation1}.
-	\item A key difference between Theorems~\ref{thm:laplacian_eigenmaps_estimation1} and~\ref{thm:laplacian_smoothing_estimation1} is that the latter holds only when $d \leq 4$.
-This restriction is unsurprising, and reflects a fundamental drawback in (penalized) least squares estimators more generally.
-The underlying phenomenon is similar to that elucidated by the seminal papers~\cite{birge1993} and~\cite{birge1998}: when the dimension $d$ is large enough, the parameter space is too large, and (penalized) least squares approaches are no longer minimax optimal.
-It is interesting to note that we see a gap emerge between the graph- (Laplacian smoothing) and continuum (smoothing spline) estimator when $d = 3$: the former continues to be minimax optimal, whereas the latter is not even well-defined.
-We should also mention that in a related context (estimation over a $d$-dimensional lattice graph), \cite{sadhanala2016} require the same restriction $d \leq 4$.
-\end{itemize}
-
-\subsection{Minimax optimal hypothesis testing}
-
-Under appropriate conditions, the Laplacian eigenmaps test statistic $T_{\LE}$ has non-trivial power over all $f_0 \in H^1(P,M)$ that are separated from $0$ in $\Leb^2(P)$ norm by at least the minimax critical radius.
-These conditions include~\ref{asmp:kernel_form} and~\ref{asmp:smooth_density}, but not~\ref{asmp:kernel_radius_estimation}.
-Instead, we will require the following scaling of the neighborhood graph radius.
-\begin{enumerate}[label=(K\arabic*)]
-	\setcounter{enumi}{2}
-	\item 
-	\label{asmp:kernel_radius_testing}
-	The radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
-	\begin{equation*}
-	n^{-1/d} \leq r(n) \leq n^{-2/(4 + d)}
-	\end{equation*}
-	for all $n$ sufficiently large.
-\end{enumerate}
-
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_testing1}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}, with $d \leq 3$. Suppose that $f_0 \in H^1(P,M)$, that
-	$P$ satisfies~\ref{asmp:bounded_density}, and that the neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_testing}.
-	Then for the Laplacian eigenmaps test statistic $T_{\LE}$ as in~\eqref{eqn:laplacian_eigenmaps_test} with $\kappa = n^{2d/(4 + d)} M^{4d/(4 + d)}$, the following statement holds true for all $n$ sufficiently large: there exist constants $C_1$ and $C_2$ which do not depend on $f_0$ such that if
-	\begin{equation}
-	\label{eqn:laplacian_eigenmaps_testing}
-	\norm{f_0}_{\Leb^2(P)}^2 \geq \frac{C_1}{\delta^2} M^{2d/(4 + d)} n^{-4/(4 + d)}
-	\end{equation}
-	then
-	\begin{equation*}
-	\Pbb \biggl(T_{\LE} \geq \frac{\kappa}{n} + \frac{1}{\delta}\sqrt{\frac{2\kappa}{n^2}}\biggr) \geq 1 - C_2\Bigl(\delta + r^{-d} \exp\{- c n r^d\}\Bigr).
-	\end{equation*}
-\end{theorem}
-\begin{itemize}
-	\item Explain why Theorem~\ref{thm:laplacian_eigenmaps_testing1} implies minimax optimality.
-	\item Include results regarding the low-smoothness regime (i.e. the case when $d \geq 4$.)
-\end{itemize}
-
-\section{Higher-order smoothness classes}
-\label{sec:higher_order_smoothness_classes}
-The estimation and testing rates derived in Section~\ref{sec:minimax_optimal_graph_Laplacian_methods} reflect a bonafide curse of dimensionality: as  the dimension $d$ gets larger, inference quickly becomes (much) harder.
-As covered in Section~\ref{sec:problem_setup_and_background}, it is well known that by making additional structural assumptions---either on the regression function $f_0$, the distribution $P$, or both---one can derive improved minimax rates.
-In the following two sections, we show that under such assumptions, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} can in certain situations obtain the faster (known) minimax rates in the nonparametric estimation and testing problems.
-
-In this section, we will assume that $f_0 \in H^s(P,M)$ for some integer $s > 1$. In other words, we will assume that $f_0$ has at least two smooth derivatives.
-
-\subsection{Minimax optimal estimation}
-\label{subsec:minimax_estimation_higher_order}
-Under the following three conditions, the estimation error of the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves the minimax optimal rate for any $f_0 \in H^s(P,M)$.
-\begin{enumerate}[label=(K\arabic*)]
-	\setcounter{enumi}{2}
-	\item 
-	\label{asmp:kernel_radius_higher_order}
-	\textbf{Neighborhood graph radius}:
-	The sequence of radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
-	\begin{equation*}
-	n^{-1/(2(s - 1) + d)}\leq r(n) \leq M^{2/(2s + d)}n^{-1/(2s + d)}
-	\end{equation*}
-	for all $n$ sufficiently large.
-\end{enumerate}
-\begin{enumerate}[label=(F\arabic*)]
-	\setcounter{enumi}{0}
-	\item 
-	\label{asmp:boundary} \textbf{Boundary condition}; The regression function $f_0$ is zero-trace. In other words, $f_0 \in H_0^s(P,M)$. 
-\end{enumerate}
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{1}
-	\item 
-	\label{asmp:smooth_density}\textbf{$\beta$-Holder density}: The design distribution $P$ has a density $p \in C^{\beta}([0,1]^d;p_{\max})$ for some $\beta \geq 0$. 
-\end{enumerate}
-
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_estimation}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}.
-Suppose that $f_0 \in H_0^s(P,M)$, that $P$ satisfies~\ref{asmp:smooth_density} with $\beta = s - 1$, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_higher_order}.
-Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as in~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2s + d)}n^{d/(2s + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
-	\begin{equation*}
-	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4s + d)} n^{-2s/(2s + d)}
-	\end{equation*}
-	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
-\end{theorem}
-
-We note that the assumption~\ref{asmp:kernel_radius_higher_order} is rather mild, since the user can choose how to construct the neighborhood graph $G_{n,r}$.
-The assumptions~\ref{asmp:boundary} and~\ref{asmp:smooth_density} are more stringent.
-That being said, some restrictions on the boundary behavior of the regression function $f_0$ are typical in the analysis of orthogonal series estimators (see e.g.
-\textcolor{red}{Tsybakov}).
-On the other hand, restrictions on the smoothness of the density $p$ are also typical when analyzing orthogonal series estimators under a random design setup (see e.g.
-\textcolor{red}{Mukherjee16}).
-
-\begin{enumerate}
-	\item \textcolor{red}{(TODO)}: Comment a bit on the proof strategy; i.e., 
-\end{enumerate}
-
-\subsection{Minimax optimal hypothesis testing}
-\label{subsec:minimax_testing_higher_order}
-
-Under appropriate conditions, the Laplacian eigenmaps test statistic $T_{\LE}$ has non-trivial power over all $f_0 \in H_0^s(P,M)$ that are separated from $0$ in $\Leb^2(P)$ norm by at least the minimax critical radius.
-These conditions include~\ref{asmp:kernel_form} and~\ref{asmp:smooth_density}.
-Additionally, we will require the following scaling of the neighborhood graph radii.
-\begin{enumerate}[label=(K\arabic*)]
-	\setcounter{enumi}{2}
-	\item 
-	\label{asmp:kernel_radius_testing}
-	The radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
-	\begin{equation*}
-	n^{-1/(2(s - 1) + d)} \leq r(n)~~\textrm{and}~~ r(n) \leq n^{-2/(4s + d)}
-	\end{equation*}
-	for all $n$ sufficiently large.
-\end{enumerate}
-In one sense this is still a mild condition, in that the neighborhood graph radius is still a tuning parameter to be selected by the user.
-However, some elementary calculations show that when $d > 4$, it is in fact the case that $n^{-1/(2(s - 1) + d)} > n^{-2/(4s + d)}$; hence no choice of $r$ can satisfy~\ref{asmp:kernel_radius_testing}.
-As a result, our theory applies only when $d \leq 4$.
-
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_testing}
-	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}. Suppose that $d \leq 4$, that $f_0 \in H_0^s(P,M)$, that
-	$P$ satisfies~\ref{asmp:smooth_density} with $\beta = s - 1$, and that the neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_testing}.
-	Then for the Laplacian eigenmaps test statistic $T_{\LE}$ given by~\eqref{eqn:laplacian_eigenmaps_test} with tuning parameter $\kappa = n^{2d/(4s + d)} M^{4d/(4s + d)}$, the following statement holds true for all $n$ sufficiently large: there exist constants $C_1$ and $C_2$ which do not depend on $f_0$ such that if
-	\begin{equation}
-	\label{eqn:laplacian_eigenmaps_testing}
-	\norm{f_0}_{\Leb^2(P)}^2 \geq \frac{C_1}{\delta^2} M^{2d/(4s + d)} n^{-4s/(4s + d)}
-	\end{equation}
-	then
-	\begin{equation*}
-	\Pbb \biggl(T_{\LE} \geq \frac{\kappa}{n} + \frac{1}{\delta}\sqrt{\frac{2\kappa}{n^2}}\biggr) \geq 1 - C_2\Bigl(\delta + r^{-d} \exp\{- c n r^d\}\Bigr).
-	\end{equation*}
-\end{theorem}
-
-\begin{itemize}
-	\item \textbf{Test based on the Laplacian smoothing test statistic}. 
-	\item \textbf{Low-smoothness regime}. Include results on minimax testing when $4s < d$.
-	\item \textbf{Head off potential criticism}.
-Acknowledge frankly that there is a regime---when $4 < d < 4s$---where minimax testing rates are known, but for which we cannot prove that either of our test statistics achieve these rates.
-\end{itemize}
-
-\section{Manifold Assumption}
-\label{sec:manifold_assumption}
-
-\section{Discussion}
-
-\begin{itemize}
-	\item \textbf{Relationships between $s$ and $d$}.
-Comment on the different types of relationships between $s$ and $d$.
-(1) Those due to fundamental upper bounds on testing; (2) those due to ``defects'' in the proposed methods, i.e.
-largeness of the parameter space for Laplacian smoothing, and (3) those (potentially) due to ``defects'' in our proof strategy.
-	\item \textbf{Proof techniques}.
-On the subject of our proof techniques, the careful reader will note that our proofs do not rely on convergence of the Laplacian eigenvectors $v_k(G_{n,r})$ towards eigenfunctions of a limiting (weighted) Laplace-Beltrami operator.
-
-	\textcolor{red}{(TODO) Flesh the following out.} This is helpful for at least the following two reasons.
-First, proving convergence of eigenvectors is harder than proving convergence of eigenvalues, a fact which is reflected in the gap between the best known rates for the two problems.
-Second, because in order for our methods to be minimax optimal, orthogonal series estimators rely on projecting onto $\kappa$ basis functions, where $\kappa$ is a number growing in $n$.
-Thus we have an accumulation of error problem.
-	\item \textbf{Possible extensions}.
-\end{itemize}
-
-\section{Simulations}
-\begin{itemize}
-	\item \textbf{Estimation MSE 1.} Plot $L^2(P_n)$ error as a function of $n$ and choice $f_0^{(n)}$ when $P$ is (close to) uniform, showing that Laplacian methods are competitive with their non-graph-based brethren.
-	\item \textbf{Estimation fitted values.} Show fits for two regression problems, one where the distribution $P$ is assumed to be close to uniform, one where $P$ is dramatically non-uniform.
-	\item \textbf{Estimation MSE 2.} Plot $L^2(P_n)$ error as a function of the distribution $P$, as we vary $P$ from being uniform to being very non-uniform.
-
-	\item \textbf{Testing power.} 
-\end{itemize}
-
-\begin{figure}
-	\centering
-	\includegraphics[width = .48\textwidth]{plots/graph_regression/mean_squared_errors/sobolev_MSE_by_density}
-	\caption{}
-	\label{fig:piecewise_cosine}
-\end{figure}
-
-\clearpage
-
-\appendix
-
-Many of our results follow the same general two-part proof strategy.
-First, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to functionals of the graph $G$.
-We then analyze the behavior of these functionals with respect to the particular neighborhood graph $G_{n,r}$ and give high probability (upper or lower) bounds on these functionals.
-It is in this second step where we invoke our various assumptions on the distribution $P$ and regression function $f_0$.
-
-\section{Graph-dependent error bounds}
-\label{sec:fixed_graph_error_bounds}
-
-Suppose we observe a fixed graph $G = \bigl([n],E\bigr)$ with Laplacian $\Lap_G$ and responses
-\begin{equation}
-\label{eqn:fixed_graph_regression_model}
-Y_i = f_{0,i} + \varepsilon_i
-\end{equation}
-where $f_0 = (f_{0,1},\ldots,f_{0,n}) \in \Reals^n$, and the noise variables $\varepsilon_i$ are independent $N(0,1)$. The Laplacian eigenmaps estimator of $f_0$ on $G$ is then
-\begin{equation}
-\label{eqn:le_G}
-\wh{f}_{\LE}(G) := \sum_{k = 1}^{\kappa} \biggl(\frac{1}{n}\sum_{i = 1}^{n} Y_i v_{k,i}(G) \biggr) v_k(G),
-\end{equation}
-where $(\lambda_1(G),v_1(G)),\ldots,(\lambda_n(G),v_n(G))$ are the eigenvalue/eigenvector pairs of $\Lap_G$. The Laplacian smoothing estimator of $f_0$ on $G$ is
-\begin{equation}
-\label{eqn:ls_G}
-\wt{f}_{\LS}(G) := \argmin_{f \in \Reals^n} \biggl\{ \sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_G^{s}  f \biggr\} = (\rho \Lap_G^s + \Id)^{-1}Y.
-\end{equation}
-
-In this fixed design context, the hypothesis testing problem becomes
-\begin{equation*}
-\mathbf{H}_0: f_{0} = (0,...,0) ~~\textrm{versus}~~ \mathbf{H}_a: f_{0} \neq (0,...,0)
-\end{equation*}
-and two test statistics for this purpose are
-\begin{equation}
-\label{eqn:le_ts_G}
-T_{\LE}(G) := \frac{1}{n} \Bigl\|\wh{f}_{\LE}(G)\Bigr\|_2^2 
-\end{equation}
-and
-\begin{equation}
-\label{eqn:ls_ts_G}
-T_{\LS}(G) := \frac{1}{n} \Bigl\|\wt{f}_{\LS}(G)\Bigr\|_2^2 
-\end{equation}
-We note that~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} specialize~\eqref{eqn:le_G}-\eqref{eqn:ls_ts_G} to the case where $G$ = $G_{n,r}$.
-
-\subsection{Error bounds for linear smoothers}
-Let $S \in \Reals^{n \times n}$ be a square, symmetric matrix, and let $\check{f} = SY$ be a linear estimator of $f_0$.
-Both~\eqref{eqn:le_G} and~\eqref{eqn:ls_G} can be written in this form, and Lemma~\ref{lem:linear_smoother_fixed_graph_estimation} is therefore useful in analyzing them.
-Let $\lambdavec(S) = (\lambda_1(S),\ldots,\lambda_n(S)) \in \Reals^n$ denote the eigenvalues of $S$ and let $v_k(S)$ denote the corresponding \emph{unit-norm} eigenvectors, so that $S = \sum_{k = 1}^{n} \lambda_k(S) \cdot v_k(S) v_k(S)^T$.
-Denote $Z_k = v_k(S)^T \varepsilon$, and observe that $Z = (Z_1,\ldots,Z_n) \sim N(0,\Id)$.
-
-\begin{lemma}
-	\label{lem:linear_smoother_fixed_graph_estimation}
-	Let $\check{f} = SY$ for a square, symmetric matrix, $S \in \Reals^{n \times n}$ satisfying $\lambda_n(S) \leq 1$. Then
-	\begin{equation*}
-	\Pbb_{f_0}\biggl(\frac{1}{n}\bigl\|\check{f} - f_0\bigr\|_2^2 \geq \frac{10}{n} \bigl\|\lambdavec(S)\bigr\|_2^2 + \frac{2}{n}\bigl\|(S - I)f_0\bigr\|_2^2\biggr) \leq 1 - \exp\Bigl(-\bigl\|\lambdavec(S)\bigr\|_2^2\Bigr)
-	\end{equation*}
-\end{lemma}
-\begin{proof}[\textbf{Proof of Lemma~\ref{lem:linear_smoother_fixed_graph_estimation}}]
-	It holds that
-	\begin{align*}
-	\frac{1}{n}\bigl\|\check{f} - f_0\bigr\|_2^2 & \leq \frac{2}{n}\Bigl(\bigl\|\check{f} - \Ebb_{f_0}[\check{f}]\bigr\|_2^2 + \bigl\|\Ebb_{f_0}[\check{f}] - f_0\bigr\|_2^2\Bigr) \\ 
-	& = \frac{2}{n}\Bigl(\bigl\|S\varepsilon\bigr\|_2^2 + \bigl\|(S - I)f_0\bigr\|_2^2\Bigr)
-	\end{align*}
-	Writing $\norm{S\varepsilon}_2^2 = \sum_{k = 1}^{n} \lambda_k(S)^2 Z_k^2$, the claim follows from the result of \textcolor{red}{Laurent and Massart} on concentration of $\chi^2$-random variables, which for completeness we restate in Lemma~\ref{lem:chi_square_bound}.
-To be explicit, taking $t = \norm{\lambdavec(S)}_2^2$ in Lemma~\ref{lem:chi_square_bound} completes the proof of Lemma~\ref{lem:le_fixed_graph_estimation}.
-\end{proof}
-
-On the testing side, the statistic $\norm{\check{f}}_2^2 = Y^T S^2 Y$ is a $U$-statistic of order $2$.
-Both~\eqref{eqn:le_ts_G} and~\eqref{eqn:ls_ts_G} can be written in this form, and Lemma~\ref{lem:linear_smoother_fixed_graph_testing} is therefore useful in analyzing them.
-\begin{lemma}
-	\label{lem:linear_smoother_fixed_graph_testing}
-	Let $T = Y^T S^2 Y$ for a square, symmetric matrix $S \in \Reals^{n \times n}$. Define the threshold $t_b$ to be 
-	\begin{equation}
-	t_{b} := \norm{\lambdavec(S)}_2^2 + 2b \norm{\lambdavec(S)}_4^2
-	\end{equation}
-	Suppose the eigenvalues $0 \leq \lambda_{1}(S) \leq \lambda_{n}(S) \leq 1$. Then,
-	\begin{itemize}
-		\item \textbf{Type I error.}
-		\begin{equation}
-		\label{eqn:linear_smoother_fixed_graph_testing_typeI}
-		\Pbb_0\bigl(T > t_b\bigr) \leq \frac{1}{b^2}
-		\end{equation}
-		\item \textbf{Type II error.} Suppose further that
-		\begin{equation}
-		\label{eqn:linear_smoother_fixed_graph_testing_critical_radius}
-		f_0^T S^2 f_0 \geq 4b \norm{\lambdavec(S)}_4^2.
-		\end{equation}
-		Then
-		\begin{equation}
-		\label{eqn:linear_smoother_fixed_graph_testing_typeII}
-		\Pbb_{f_0}\bigl(T \leq t_b\bigr) \leq \frac{4}{b^2} + \frac{8}{b \norm{\lambdavec(S)}_4^{2}} 
-		\end{equation}
-	\end{itemize}
-\end{lemma}
-\begin{proof}[\textbf{Proof of Lemma~\ref{lem:linear_smoother_fixed_graph_testing}}]
-	We compute the mean and variance of $T$ as a function of $f_0$, then apply Chebyshev's inequality.
-
-	\textit{Mean.} Writing $Y = f_0 + \varepsilon$, we make use of the eigendecomposition of $S$ to obtain
-	\begin{equation}
-	\label{pf:linear_smoother_fixed_graph_testing1}
-	\begin{aligned}
-	T & = f_0^T S^2 f_0 + 2 f_0^T S^2 \varepsilon + \varepsilon^T S^2 \varepsilon \\
-	& = f_0^T S^2 f_0 + 2 f_0^T S^2 \varepsilon + \sum_{k = 1}^{n}  \bigl(\lambda_k(S)\bigr)^2 (\varepsilon^T v_k(S))^2 \\
-	& = f_0^T S^2 f_0 + 2 f_0^T S^2 \varepsilon + \sum_{k = 1}^{n}  \bigl(\lambda_k(S)\bigr)^2 Z_k^2,
-	\end{aligned}
-	\end{equation}
-	implying
-	\begin{equation}
-	\label{pf:linear_smoother_fixed_graph_testing_mean}
-	\Ebb_{f_0}[T] = f_0^T S^2 f_0 + \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^2.
-	\end{equation}
-
-	\textit{Variance.} Starting from~\eqref{pf:linear_smoother_fixed_graph_testing1} and recalling the basic fact $\Var(Z_k^2) = 2$, we derive
-	\begin{equation}
-	\label{pf:linear_smoother_fixed_graph_testing_var}
-	\Var_{f_0}[T] \leq 8 f_0^T S^4 f_0 + 4 \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^4 \leq 8 f_0^T S^2 f_0 + 4 \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^2
-	\end{equation}
-	where the second inequality follows since by assumption $\lambda_{n}(S) \leq 1$.
-
-	\textit{Bounding Type I and Type II error.} The bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeI} follows directly from Chebyshev's inequality, along with our above calculations on the mean and variance of $T$.
-
-	The bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeII} also follows from Chebyshev's inequality, as can be seen by the following manipulations,
-	\begin{equation*}
-	\begin{aligned}
-	\Pbb_{f_0}\bigl(T \leq t_b\bigr) & = \Pbb_{f_0}\bigl(T - \Ebb_{f_0}[T] \leq t_b - \Ebb_{f_0}[T]\bigr) \\
-	& \overset{(i)}{\leq} \Pbb_{f_0}\bigl(\abs{T - \Ebb_{f_0}[T]} \geq \abs{t_b - \Ebb_{f_0}[T]}\bigr) \\ 
-	& \overset{(ii)}{\leq} 4 \frac{\Var_{f_0}[T]}{(f_0^T S^2 f_0)^2} \\
-	& \overset{(iii)}{\leq} \frac{32}{f_0^T S^2 f_0} + \frac{4}{b^2} \\
-	& \overset{(iv)}{\leq} \frac{8}{b} \Biggl(\sum_{k = 1}^{n}\bigl(\lambda_k(S)\bigr)^4\Biggr)^{-1/2} + \frac{4}{b^2}
-	\end{aligned}
-	\end{equation*}
-	In the previous expression, $(i)$ and $(ii)$ follow since assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and equation~\eqref{pf:linear_smoother_fixed_graph_testing_mean} together imply $\Ebb_{f_0}(T) - \frac{1}{2}f_0^T S^2f_0 \geq t_b$, $(iii)$ follows from assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and the inequality~\eqref{pf:linear_smoother_fixed_graph_testing_var}, and $(iv)$ follows assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}.
-\end{proof}
-\subsection{Analysis of Laplacian eigenmaps}
-
-In Lemma~\ref{lem:le_fixed_graph_estimation}, we upper bound the mean squared error of $\wh{f}_{\LE}(G)$.
-Our bound will be stated as a function of $f_0^T \Lap_G^s f_0$--a measure of the smoothness the signal $f_0$ displays over the graph $G$--as well as the $\kappa$th eigenvalue $\lambda_{\kappa}(G)$.
-
-\begin{lemma}
-	\label{lem:le_fixed_graph_estimation}
-	For any integer $1 \leq \kappa \leq n$ and any $s > 0$,
-	\begin{equation}
-	\label{eqn:le_fixed_graph_estimation}
-	\frac{1}{n}\bigl\|\wh{f}_{\LE}(G) - f_0\bigr\|_2^2 \leq 2\frac{f_0^T \Lap_{G}^s f_0}{n\lambda_{\kappa}(G)^s} + \frac{10\kappa}{n}
-	\end{equation}
-	with probability at least $1 - \exp(-\kappa)$.
-\end{lemma}
-\begin{proof}[\textbf{Proof of Lemma~\ref{lem:le_fixed_graph_estimation}}]
-	Note that if $\lambda_{\kappa}(G) = 0$ the claim is trivial, and we therefore only cover the case when $\lambda_{\kappa}(G) > 0$.
-
-	Letting $\wh{S} = \frac{1}{n}\sum_{k = 1}^{n} v_k(G) v_k(G)^T$, the estimator $\wh{f}_{\LE}(G) = \wh{S}Y$.
-The matrix $\wh{S} \in \Reals^{n \times n}$ is symmetric, and satisfies $\lambda_k(\wh{S}) \in \{0,1\}$ for all $k \in [n]$.
-The claim then follows from Lemma~\ref{lem:le_fixed_graph_estimation} upon noting first that $\norm{\lambdavec(\wh{S})}_2^2 = \kappa$, and second that
-	\begin{equation}
-	\label{pf:le_fixed_graph_estimation_1}
-	\begin{aligned}
-	\bigl\|(\wh{S} - I)f_0\bigr\|_2^2 & = \sum_{k = \kappa + 1}^{n} \bigl(f_0^T v_k(G)\bigr)^2 \\
-	& \leq \frac{1}{\lambda_{\kappa}(G)^s}\sum_{k = \kappa + 1}^{n} \lambda_{k}(G)^s \bigl(f_0^T v_k(G)\bigr)^2 \\
-	& \leq \frac{1}{\lambda_{\kappa}(G)^s}\sum_{k = 1}^{n} \lambda_{k}(G)^s \bigl(f_0^T v_k(G)\bigr)^2 \\
-	& = \frac{f_0^T \Lap_G^s f_0}{\lambda_{\kappa}(G)^s}.
-	\end{aligned}
-	\end{equation} 
-\end{proof}
-
-In Lemma~\ref{lem:le_fixed_graph_testing}, we upper bound the testing error of a test based on $T_{\LE}(G)$.
-
-\begin{lemma}
-	\label{lem:le_fixed_graph_testing}
-	Define the threshold $\wh{t}_b$ to be
-	\begin{equation*}
-	\wh{t}_b := \frac{\kappa}{n} + \frac{2b\sqrt{\kappa}}{n}
-	\end{equation*}
-	Then each of the following hold for any $\kappa \in [n]$ and any $b \geq 1$,
-	\begin{enumerate}
-		\item \textbf{Type I error:} Under the null hypothesis $f_0 = 0$,
-		\begin{equation}
-		\label{eqn:graph_spectral_type_I_error}
-		\Pbb_{f_0}\biggl(T_{\LE}(G) > \wh{t}_b\biggr) \leq \frac{1}{b^2}.
-		\end{equation}
-		\item \textbf{Type II error:} Suppose that
-		\begin{equation}
-		\label{eqn:le_fixed_graph_testing_critical_radius}
-		\frac{1}{n} \norm{f_0}_2^2 \geq \frac{f_0^T \Lap_G^s f_0}{n \lambda_{\kappa}(G)^s} + \frac{4b\sqrt{\kappa}}{n}
-		\end{equation}
-		Then,
-		\begin{equation}
-		\label{eqn:graph_spectral_type_II_error}
-		\Pbb_{f_0}\biggl(T_{\LE}(G) \leq \wh{t}_b\biggr) \leq \frac{4}{b^2} + \frac{8}{b\sqrt{\kappa}}.
-		\end{equation}
-	\end{enumerate}
-\end{lemma}
-\begin{proof}[\textbf{Proof of Lemma~\ref{lem:le_fixed_graph_testing}}]
-	Letting $\wh{S} = \frac{1}{n}\sum_{k = 1}^{\kappa} v_k(G) v_k(G)^T$, our test statistic $T_{\LE}(G) = \frac{1}{n} Y^T \wh{S}^2 Y$.
-The matrix $\wh{S}$ is symmetric and positive semidefinite, with eigenvalues $\lambda_k(\wh{S}) \in \{0,1\}$ for all $k \in [n]$, so $\norm{\lambdavec(\wh{S})}_2^2 = \norm{\lambdavec(\wh{S})}_4^4 = \kappa$ and clearly $n \wh{t}_b = \norm{\lambdavec(\wh{S})}_2^2 + 2b \norm{\lambdavec(\wh{S})}_4^2$.
-The desired result thus follows from Lemma~\ref{lem:linear_smoother_fixed_graph_testing}.
-To see that the conditions of this Lemma are met, note first that $\lambda_k(\wh{S}) \in \{0,1\}$ for all $k \in [n]$, and so it remains only to verify \eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}.
-Invoking assumption~\eqref{eqn:le_fixed_graph_testing_critical_radius} along with the inequality~\eqref{pf:le_fixed_graph_estimation_1}, we obtain
-	\begin{align*}
-	f_0^T \wh{S}^2 f_0 & = \norm{f_0}_2^2 - \sum_{k = \kappa + 1}^{n} \bigl(f_0^T v_k(G)\bigr)^2 \\
-	& \geq \frac{f_0^T \Lap_G^s f_0}{\lambda_{\kappa}(G)^s} + 4b\sqrt{\kappa} - \sum_{k = \kappa + 1}^{n} \bigl(f_0^T v_k(G)\bigr)^2 \\
-	& \geq 4b\sqrt{\kappa}.
-	\end{align*}
-	Reiterating that $\norm{f_0}_4^2 = \sqrt{\kappa}$, we see that~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} is met, completing the proof.
-\end{proof}
-
-\subsection{Analysis of Laplacian Smoothing}
-In Lemma~\ref{lem:ls_fixed_graph_estimation}, we upper bound the mean squared error of $\wt{f}_{\LS}(G)$.
-\begin{lemma}
-	\label{lem:ls_fixed_graph_estimation}
-	For any $\rho,s > 0$,
-	\begin{equation}
-	\label{eqn:ls_fixed_graph_estimation_prob}
-	\frac{1}{n}\bigl\|\wt{f}_{\LS}(G) - f_0\bigr\|_2^2 \leq \frac{2\rho}{n} \bigl(f_0^T \Lap_{G}^s f_0\bigr) + \frac{10}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k}(G)^s + 1\bigr)^2}
-	\end{equation}
-	with probability at least $1 - \exp\Bigl(-\sum_{k = 1}^{n}\bigl(\rho \lambda_k(G)^s + 1\bigr)^{-2}\Bigr)$.
-\end{lemma}
-\begin{proof}[\textbf{Proof of Lemma~\ref{lem:ls_fixed_graph_estimation}}]
-	Letting $\wt{S} = (\Id + \rho \Lap_G^s)^{-1}$, the estimator $\wt{f}_{\LS}(G) = \wt{S}Y$.
-The matrix $\wt{S} \in \Reals^{n \times n}$ is symmetric, and satisfies $\lambda_k(\wt{S}) \in (0,1)$ for all $k \in [n]$.
-The claim then follows from Lemma~\ref{lem:linear_smoother_fixed_graph_estimation} upon noting first that
-	\begin{equation*}
-	\bigl\|\lambdavec(\wt{S})\bigr\|_2^2 = \sum_{k = 1}^{n} \frac{1}{\bigl(1 + \rho \lambda_k(G)^s\bigr)^2}
-	\end{equation*} 
-	and second that
-	\begin{equation*}
-	\begin{aligned}
-	\bigl\|(\wt{S} - I)f_0\bigr\|_2^2 & = f_0^T \Lap_G^{s/2} \Lap_G^{-s/2}\bigl(\wt{S} - \Id\bigr) \Lap_G^{-s/2} \Lap_G^{s/2} f_0 \\
-	& \leq f_0^T \Lap_G^s f_0 \cdot \lambda_n\Bigl(\Lap_G^{-s/2}\bigl(\wt{S} - \Id\bigr)\Lap_G^{-s/2}\Bigr) \\
-	& = f_0^T \Lap_G^s f_0 \cdot \max_{k \in [n]} \biggl\{\frac{1}{\lambda_k(G)^s} \Bigl(1 - \frac{1}{\rho\lambda_k(G)^s + 1}\Bigr) \biggr\} \\
-	& = f_0^T \Lap_G^s f_0 \cdot \rho.
-	\end{aligned}
-	\end{equation*} 
-	where we write $\Lap_G^{-1}$ for the pseudoinverse of $\Lap_G$.
-\end{proof}
-
-In Lemma~\ref{lem:ls_fixed_graph_testing}., we upper bound the testing error of a test based on $T_{\LS}(G)$.
-\begin{lemma}
-	\label{lem:ls_fixed_graph_testing}
-	Define the threshold $\wt{t}_b$ to be
-	\begin{equation*}
-	\wt{t}_b := \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k(G)^s + 1\bigr)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k(G)^s + 1\bigr)^4}}
-	\end{equation*}
-	Then each of the following holds for any $\rho,s > 0$ and any $b \geq 1$.
-	\begin{itemize}
-		\item \textbf{Type I error.}
-		\begin{equation}
-		\label{eqn:ls_fixed_graph_testing_typeI}
-		\Pbb_0\Bigl(T_{\LS}(G) > \wt{t}_b\Bigr) \leq \frac{1}{b^2}
-		\end{equation}
-		\item \textbf{Type II error.} If
-		\begin{equation}
-		\label{eqn:ls_fixed_graph_testing_critical_radius}
-		\frac{1}{n}\norm{f_0}_2^2 \geq \frac{2 \rho}{n} \bigl(f_0^T \Lap_G^s f_0\bigr) + \frac{4b}{n} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G)^s + 1)^4} \Biggr)^{1/2}
-		\end{equation}
-		then
-		\begin{equation}
-		\label{eqn:ls_fixed_graph_testing_typeII}
-		\Pbb_{f_0}\Bigl(T_{\LS}(G) \leq \wt{t}_b\Bigr) \leq \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G)^s + 1)^4} \Biggr)^{-1/2}
-		\end{equation}
-	\end{itemize}
-\end{lemma}
-
-\begin{proof}[\textbf{Proof of Lemma~\ref{lem:ls_fixed_graph_testing}}]
-	Let $\wt{S} := (\rho \Lap^s + \Id)^{-1}$.
-The matrix $\wt{S} \in \Reals^{n \times n}$ is symmetric and positive semidefinite, and our test statistic $T_{\LS}(G) = \frac{1}{n}Y^T \wt{S}^2 Y$.
-The desired result thus follows from Lemma~\ref{lem:linear_smoother_fixed_graph_testing}.
-To see that the conditions of this Lemma are satisfied, we first note that since
-	\begin{equation*}
-	\lambda_k(\wt{S}) = \frac{1}{(\rho\lambda_k(G)^s + 1)}
-	\end{equation*}
-	and $\rho, \lambda_k(G) > 0$, it is evident that $\lambda_{n}(\wt{S}) \leq 1$.  Then, by assumption~\eqref{eqn:ls_fixed_graph_testing_critical_radius}
-	\begin{equation*}
-	f_0^{T} \wt{S}^2 f_0 = \norm{f_0}_2^2 - f_0^T(I - \wt{S}^2)f_0 \geq 2 \rho (f_0^T \Lap^s f_0) + f_0^T(I - \wt{S}^2)f_0 + 4b \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G)^s + 1)^4} \Biggr)^{-1/2},
-	\end{equation*}
-	and along with the following calculations,
-	\begin{equation*}
-	\begin{aligned}
-	f_0^T \Bigl(\Id - \wt{S}^2\Bigr) f_0  & = f_0^T L^{s/2} L^{-s/2}\Bigl(\Id - \wt{S}^2\Bigr) L^{-s/2} L^{s/2} f_0 \\ 
-	& \leq f_0^T L^{s} f_0 \cdot  \lambda_{\max}\biggl(L^{-s/2}\Bigl(\Id - \wt{S}^2\Bigr) L^{-s/2}\biggr) \\ 
-	& \overset{(i)}{=}  f_0^T L^{s} f_0 \cdot \max_{k} \biggl\{ \frac{1}{\lambda_k(G)^s} \Bigl(1 - \frac{1}{(\rho \lambda_k(G)^s + 1)^2}\Bigr) \biggr\} \\
-	& \overset{(ii)}{\leq} f_0^T L^{s} f_0 \cdot 2\rho,
-	\end{aligned}
-	\end{equation*}
-	we have that
-	\begin{equation*}
-	f_0^{T} \wt{S}^2 f_0 \geq 2b \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k(G)^s + 1)^4} \biggr)^{-1/2}.
-	\end{equation*} 
-	In other words condition~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} in Lemma~\ref{lem:linear_smoother_fixed_graph_testing} is met, and applying that Lemma completes the proof.
-
-	(In the previous derivation: in $(i)$ the maximum is over all indices $k$ such that the eigenvalue $\lambda_k(G)$ is strictly positive; and $(ii)$ follows from the basic algebraic identity $1 - 1/(1 + \rho x)^2 \leq 2 \rho x$ for any $x, \rho > 0$.
-\end{proof}
-
-\clearpage
-To prove Theorems~\ref{thm:laplacian_eigenmaps_estimation},\textcolor{red}{...}, we would like to apply Lemmas~\ref{lem:fixed_graph_estimation},\textcolor{red}{...} in the random design context, i.e.
-to the random geometric graph $G_{n,r}$.
-To do so, we must establish the following (high probability) bounds: (i) an upper bound on the graph Sobolev seminorm of the regression function $f_0 \Lap_{n,r}^s f_0$, (ii) a lower bound on the Laplacian eigenvalue $\lambda_{\kappa}(G_{n,r})$ as a function of $\kappa$, and (iii) a lower bound on the empirical norm $\norm{f}_{n}$ as a function of $\norm{f}_{L^2(P)}$.
-
-\section{Bounds on the graph Sobolev semi-norm}
-\label{sec:graph_sobolev_seminorm}
-We first bound $f_0^T \Lap_{n,r}^s f_0$ in expectation.
-\begin{lemma}
-	\label{lem:roughness_functional_expectation_sobolev}
-	Let $\Xset$ be a Lipschitz domain and let $s \geq 1$ be an integer.
-Suppose that \textcolor{red}{$f \in H_0^{s}(P,M)$}, and that $P$ admits a density $p \in C^{s-1}(\Xset,p_{\max})$ for some constant $p_{\max} < \infty$.
-Let $\{r(n)\}$ be a sequence of connectivity radii such that $n^{-1/(2(s - 1) + d)} \leq r(n)$ for all $n$ sufficiently large, and $r(n) \to 0$ as $n \to \infty$.
-Then \textcolor{red}{for any $2$nd-order kernel $K$}, there exists a constant $c$ depending only on $p_{\max}, K_{\max}$, $s$, $d$ and $\Xset$ such that the resulting neighborhood graph Laplacian satisfies
-	\begin{equation}
-	\label{eqn:roughness_functional_expectation_sobolev}
-	\Ebb\bigl[f^T L_{n,r}^s f\bigr] \leq c \cdot n \cdot (nr^{d+2})^s \cdot M^2
-	\end{equation}
-	for all $n$ sufficiently large.
-\end{lemma}
-
-\section{Bounds on graph eigenvalues}
-\label{sec:graph_eigenvalues}
-Lemma~\ref{lem:neighborhood_eigenvalue} establishes a scaling rate which holds for all eigenvalues $\lambda_{k}(G_{n,r})$, assuming appropriate regularity conditions on $P$.
-\begin{lemma}
-	\label{lem:neighborhood_eigenvalue}
-	Suppose the distribution $P$ satisfies~\ref{asmp:bounded_density}.
-Then there exists constants $c$ and $C$ which depend only on $p_{\min}, p_{\max}$ and $d$, such that with probability at least $1 - C r^{-d} \exp\bigl\{- c n r^d\bigr\}$, for all $2 \leq k \leq n$
-	\begin{equation}
-	\label{eqn:neighborhood_eigenvalue}
-	\lambda_k(G_{n,r}) \geq c \cdot \min\Bigl\{nr^{d + 2}k^{2/d}, nr^d\Bigr\}.
-	\end{equation}	
-\end{lemma}
-As those readers familiar with Sobolev spaces will be aware of, the scaling $k^{2/d}$ is the \textcolor{red}{typical scaling of a Sobolev ellipsoid}, and is thus the scaling of the eigenvalues we desire.
-On the other hand, the second term on the right hand side of~\eqref{eqn:neighborhood_eigenvalue} is an artifact of the well known inequality $\lambda_n(G) \leq 2 \max_{i = 1,\ldots,n} \mathbf{D}_{ii}$--- in words, that the maximum eigenvalue of the Laplacian can never be more than twice the maximum degree of the graph.
-Doing some algebra, we may deduce the conditions on $k$ and $r$ under which the graph eigenvalue follows the desired scaling, i.e.
-$\lambda_k(G_{n,r}) \sim n r^{d + 2}k^{2/d}$.
-
-First, when $k = M^{2/(2s + d)} n^{d/(2s + d)}$, it holds that for any $r \leq M^{-2/(d(2s +d))} n^{-1/(2s + d)}$, 
-\begin{equation}
-\label{eqn:neighborhood_eigenvalue_1}
-\lambda_{k}(G_{n,r}) \geq c \cdot nr^{d+2}{k}^{2/d}
-\end{equation}
-with probability at least $1 - C r^{-d} \exp\{- c n r^d\}$. Second, when $k$
-
-\section{Proofs of Theorems}
-\label{sec:proofs_theorems}
-With the results of Sections~\ref{sec:fixed_graph_error_bounds}-\ref{sec:graph_eigenvalues} established, the proofs of our main theorems are quite straightforward.
-
-\subsection{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation}}
-\label{subsec:laplacian_eigenmaps_estimation_pf}
-
-\paragraph{\eqref{eqn:fixed_graph_estimation_sobolev_rate_2} holds with high probability.}
-
-Since $\Lap_{n,r}$ is a positive semi-definite matrix, $f_0^T \Lap_{{n,r}}^s f_0$ must be non-negative. We may therefore apply Markov's inequality and conclude
-\begin{equation}
-\label{eqn:roughness_functional_probability_sobolev}
-f_0^T \Lap_{{n,r}}^s f_0 \leq  \frac{1}{\delta} \cdot c \cdot n \cdot (nr^{d+2})^s \cdot M^2
-\end{equation}
-with probability at least $1 - \delta$.
-
-\paragraph{\eqref{eqn:fixed_graph_estimation_sobolev_rate_1} holds with high probability.}
-
-\paragraph{Putting the pieces together.}
-
-Let $n$ be sufficiently large so that $n^{-1/(2(s - 1)+ d)} \leq r(n) \leq M^{-2/(d(2s + d))} n^{-1/(2s + d)}$.
-Then by~\eqref{eqn:roughness_functional_probability_sobolev} and~\eqref{eqn:neighborhood_eigenvalue_1}, there exists a set $\mathcal{E}(f_0) \subseteq \Xset^n$ of probability at least $1 - \delta - Cr^{-d}\exp\{-cnr^d\}$ such that whenever ${\bf X} \in \mathcal{E}(f_0)$,
-\begin{equation*}
-f_0^T \Lap_{n,r}^s f_0 \leq C \cdot n \cdot C_n^s \cdot M^2
-\end{equation*}
-and
-\begin{equation*}
-\lambda_{\kappa}(G_{n,r}) \geq c \cdot nr^{d+2}\kappa^{2/d}
-\end{equation*}
-Theorem~\ref{thm:laplacian_eigenmaps_estimation} then follows by Corollary~\ref{cor:fixed_graph_estimation_sobolev_rate}.
-
-\section{Concentration Inequalities}
-\begin{lemma}
-	\label{lem:chi_square_bound}
-	Let $\xi_1,\ldots,\xi_N$ be independent $N(0,1)$ random variables, and let $U := \sum_{k = 1}^{N} a_k(\xi_k^2 - 1)$.  Then for any $t > 0$,
-	\begin{equation*}
-	\Pbb\Bigl[U \geq 2 \norm{a}_2 \sqrt{t} + 2 \norm{a}_{\infty}t\Bigr] \leq \exp(-t).
-	\end{equation*}
-	In particular if $a_k = 1$ for each $k = 1,\ldots,N$, then
-	\begin{equation*}
-	\Pbb\Bigl[U - N \geq 2\sqrt{N t} + 2t\Bigr] \leq \exp(-t).
-	\end{equation*}
-\end{lemma}
-
 \bibliographystyle{plainnat}
-\bibliography{../graph_testing_bibliography}
+\bibliography{../../graph_testing_bibliography}
 

### 2021-02-11 manuscripts/laplacian_eigenmaps/outline.tex
Another round of re-structuring.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=blue,
	citecolor=blue
}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{tikz}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}
\newcommand{\lambdavec}{\boldsymbol{\lambda}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}
\newcommand{\Lap}{{\bf L}}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}
\newcommand{\Id}{I}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\SM}{\mathrm{SM}}
\newcommand{\OS}{\mathrm{FS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Minimax-optimal regression over Sobolev Spaces with Neighborhood Graphs}
\author{Alden Green}
\date{\today}
\maketitle

\section{Introduction}
\label{sec:introduction}

\textbf{(1) Geometric graphs.} Suppose one observes data $X_1,\ldots,X_n$ sampled independently from an unknown distribution $P$, and forms a geometric graph $G$---with edges $e_{ij}$ corresponding to proximity between samples $X_i$ and $X_j$---over the observed data. Geometric graphs encode information about $P$ in an extremely general manner, and have thus been leveraged to conduct many different fundamental statistical tasks. These include clustering, semi-supervised learning, classification, and regression---both estimation and goodness-of-fit testing. Though much theoretical work has been done on the consistency of graph-based learning methods---and more recently, rates of convergence have been established for some problems---little is known so far about their optimality, even for classic statistical tasks.

\textbf{(2) What we study: Regression and Laplacian eigenmaps.} In this paper we focus on regression, where in addition to the design points $X_1,\ldots,X_n$ one observes real-valued responses $Y_1,\ldots,Y_n$, and seeks to infer the regression function $f_0(x) := E[Y|X = x]$. The method we will study is \emph{Laplacian eigenmaps}, first introduced by \textcolor{red}{(Belkin 2003)}, which constructs an estimator $\wh{f}$ using eigenvectors of the graph Laplacian $L$ as features in a least-squares problem. The graph Laplacian $L$ is a difference operator, acting on functions $f: X \to \Reals$, 
\begin{equation}
\label{eqn:graph_laplacian}
Lf(X_i) = \sum_{j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)e_{ij},
\end{equation}
The Laplacian $L$ can be seen as a discretization of the Laplace-Beltrami operator \textcolor{red}{(...)}. The eigenvectors of $L$ are an orthonormal basis of $L^2(X)$, which can seen as estimates of the $L^2(P)$ orthonormal basis \textcolor{red}{(...)}. Their corresponding eigenvalues  are estimates of~\textcolor{red}{(...)}, and provide a notion of smoothness for each eigenvector---the smaller the eigenvalue, the smoother the eigenvector.

\textbf{(3) Why we study.} Laplacian eigenmaps is thus a type of \emph{spectral series} estimator. Spectral series estimators are one of the most classical ways to perform non-parametric regression. Traditional spectral series estimators use an \emph{a priori defined basis} \textcolor{red}{(...)}, and are typically defined with respect to some reference measure $Q$. In contrast, eigenvectors of the graph Laplacian are data-dependent, and adapt to the geometry of the distribution $P$ in a rich manner.

The study of (classical) spectral series estimators is rich, and their theoretical optimality is well-understood \textcolor{red}{(references)}. However, in practice these estimators suffer from some serious drawbacks.

\textbf{(4) Our contributions.} Laplacian eigenmaps avoid these drawbacks \textcolor{red}{(...)} On the other hand their statistical properties are not as well understood \textcolor{red}{(...)}. The primary contribution of our paper is to fill this theoretical gap, by answering the following question:

\begin{quote}
	\textcolor{red}{(TODO)}
\end{quote}

We show that under various classical assumptions on the regression function $f_0$ and design distribution $P$, Laplacian eigenmaps is statistically minimax optimal. \textcolor{red}{(TODO): Add a richer summary of your major results here.}

\textbf{(5) Organization.}

\section{Setup and Background}
\label{sec:setup_main_results}

In this section, we give a precise setup of the non-parametric regression problems and methods we will subsequently consider, review the minimax rates for non-parametric regression over Sobolev spaces, and summarize our main results.

\subsection{Non-parametric regression with Laplacian Eigenmaps}
\label{sec:regression_laplacian_eigenmaps}

\textbf{(1) Formal setup.}
Throughout, we assume that we observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_N$ are drawn i.i.d from a distribution $P$ supported on a compact set $\mc{X} \subseteq \Rd$, and 
\begin{equation*}
Y_i = f_0(X_i) + \varepsilon_i,
\end{equation*}
with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise with unit variance.

\textbf{(2) Laplacian eigenmaps.}

\subsection{Sobolev Classes}
\label{sec:sobolev}
We now review the Sobolev classes $H^s(\mc{X})$.

\textbf{(1) Sobolev norms, semi-norms, and balls.}

\textbf{(2) Zero-trace conditions.}

\subsection{Minimax Rates}
\label{subsec:minimax_rates_sobolev}

\textbf{(1) Estimation rates.}

\textbf{(2) Testing rates.}

\textbf{(3) Conditions under which spectral series methods achieve these rates.}

\textbf{(4) Comparison with the conditions necessary for Laplacian Eigenmaps.}

\subsection{Our contributions}
The following items summarize our major results.
\begin{itemize}
	\item \textbf{Minimax optimal estimation.} With high probability, $\norm{\wh{f} - f_0}_{n}^2 \lesssim n^{-2/(2 + d)}$ over all $f_0 \in H^1(\Xset;M)$. 
	\item \textbf{Minimax optimal testing.}
	A test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in H^1(\Xset;M)$ satisfies $\norm{f_0}_{\Leb^2(\Xset)} \gtrsim n^{-2/(4 + d)}$ and $d < 4$.
	\item \textbf{Taking advantage of higher-order smoothness conditions.} When the regression function $f_0 \in H^s(\Xset)$ and additionally the density $p \in C^{s - 1}(\mc{X})$, with high probability the error $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2s/(2s + d)}$ over all $f_0 \in H_0^s(\Xset;M)$. Additionally, a test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in H_0^s(\Xset;M)$ satisfies $\norm{f_0}_{\Leb^2(\Xset)} \gtrsim n^{-2s/(4s + d)}$ and $d < 4$.
	\item \textbf{Manifold adaptivity.}
	If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$, each of the aforementioned rates hold with $d$ replaced by $m$.
\end{itemize}
\textbf{(1) Major takeaway.} Viewed as a whole, our results show that Laplacian eigenmaps retain the strong theoretical properties of classic orthogonal series approaches, while also automatically \emph{adapting} to an unknown design distribution $P$.

\subsection{Related Work}
\begin{itemize}
	\item \cite{lee2016} considers a similar estimator --- the notable difference being the normalization of the Laplacian $L$ --- but get suboptimal rates for the supervised problem. 
	\item \cite{zhou2011} consider a similar estimator, but in the semi-supervised setting. They consider the limit as the number of unlabeled points grows to infinity, in which case the estimator is simply a classic spectral series estimator, and the analysis is standard. 
\end{itemize}

\section{Minimax Optimality of Laplacian Eigenmaps}
\label{sec:minimax_optimal_laplacian_eigenmaps}

In this section, we will show that the estimator $\wh{f}$, and a test using the statistic $\wh{T}$, achieve optimal estimation and goodness-of-fit testing rates over Sobolev classes. Throughout this section, we will assume $\mc{X}$ is a \textcolor{red}{regular, $d$-dimensional domain} and $P$ has a \textcolor{red}{regular density with respect to Lebesgue measure.}

\subsection{First-order Sobolev classes}
\label{sec:first_order_sobolev_classes}
To begin, we handle the case where $f_0 \in H^1(\mc{X})$.

\textbf{(1) Laplacian eigenmaps is an optimal estimator.} Under appropriate choices of $r$ and $\kappa$, the estimator $\wh{f}$ has (in-sample) optimal risk (up to constant factors) $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2/(2 + d)}$.

\begin{theorem}
	\label{thm:laplacian_eigenmaps_estimation_fo}
	\textcolor{red}{(TODO)}
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_estimation_fo}
	\|\wh{f} - f_0\|_n^2 \lesssim n^{-2/(2 + d)}
	\end{equation}
\end{theorem}
Some remarks:
\begin{itemize}
	\item \textbf{(1.1) How to tune hyper-parameters.} In practice, one typically tunes the hyper-parameters of an algorithm by sample-splitting or (more usually) by cross-validation. However, because $\wh{f}$ is defined only in-sample such approaches cannot be straightforwardly applied to select the graph radius $r$, or number of eigenvectors $\kappa$. We discuss out-of-sample extensions of $\wh{f}$ later, in Section~\ref{sec:out_of_sample}.
	\item \textbf{(1.2) The dependence on failure probability can be improved.} The upper bound on in-sample risk given by Equation~\eqref{eqn:laplacian_eigenmaps_estimation_fo} holds with probability $1 - C\delta$, and depends on a factor of $1/\delta$. When $f \in C^1(\mc{X})$, we can show that this depends on a factor of $(1 + 1/\delta)$, and holds with probability at least $1 - C\delta^2/n$. 
\end{itemize}

\textbf{(2) Laplacian eigenmaps is an optimal test.} Consider the test $\phi = \1\{\wh{T} \geq t\}$, where $t$ is the data-dependent threshold~\textcolor{red}{(...)}. The choice of $t$ guarantees that $\phi$ is a level-$\alpha$ test. Under appropriate choices of $r$ and $\kappa$, the test $\phi$ has non-negligible power (up to constant factors) against alternatives separated the null by at least $\|f_0\|_{\Leb^2(\mc{X})} \gtrsim n^{-4/(4 + d)}$.

\begin{theorem}
	\label{thm:laplacian_eigenmaps_testing_fo}
	\textcolor{red}{(TODO)}
\end{theorem}
Some remarks:
\begin{itemize}
	\item \textbf{(2.1) When $d \geq 4$, the testing problem is different.} When $d \geq 4$, it is not the case that $\Leb^4(\mc{X})$ continuously embeds into $H^1(\mc{X})$, and testing rates over $H^1(\mc{X})$ are not known. On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then simply taking $T = \|Y\|_n^2$ achieves the optimal rates.
	\item \textbf{(2.2) Both of the remarks after Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} also apply to testing.}
\end{itemize}

\subsection{Higher-order Sobolev classes}
\label{sec:higher_order_sobolev_classes}
In this section, we handle the case where $f_0 \in H_0^s(\mc{X})$.

\textbf{(1) Laplacian eigenmaps is an optimal estimator.}  Under appropriate choices of $r$ and $\kappa$, the estimator $\wh{f}$ has (in-sample) optimal risk (up to constant factors) $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2s/(2s + d)}$. 
\begin{theorem}
	\label{thm:laplacian_eigenmaps_estimation_ho}
	Under appropriate conditions, with probability at least $1 - \delta$,
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_estimation_ho}
	\|\wh{f} - f_0\|_n^2 \lesssim n^{-2s/(2s + d)}
	\end{equation}
\end{theorem}
Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that Laplacian eigenmaps is a rate-optimal estimator over Sobolev classes, for all scalings of $s$ and $d$. Some other remarks:
\begin{itemize}
	\item \textbf{(1.1) Optimal rates, no RKHS.} It is worth pointing out that we do not require that $2s > d$, a condition often seen in the literature. When $2s \leq d$, the Sobolev space $H^s(\mc{X})$ is quite irregular. For instance, it is not a Reproducing Kernel Hilbert Space, nor does it embed into any H\"{o}lder space. This renders certain versions of the nonparametric regression problem ill-posed---for instance when the design points $X$ are assumed to be fixed, or loss is measured in $\Leb^{\infty}$ norm. It also takes many estimators ``off the table'', most notably including RKHS-based methods such as smoothing splines. However, at these scales traditional spectral series estimators still obtain the optimal rates for random design regression measured in $\Leb^2$ norm. Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} shows that the same is true with respect to Laplacian eigenmaps.
	\item \textbf{(1.2) The zero-trace condition.} The zero-trace condition can likely be weakened. However, requiring some boundary condition on $f_0$ and its first through $(k -1 )$th-order derivatives is unavoidable. \textcolor{red}{Explain why: graphs are implicitly boundaryless objects, and the eigenvectors of $L$ converge to those eigenvectors of $\Delta_P$ which satisfy boundary conditions. Repeat that this is a standard assumption made when analyzing spectral series estimators}.
	\item \textbf{(1.3) The bounded norm and smooth density conditions.} We require that $f_0$ be bounded in the Sobolev norm $\|f_0\|_{H^s(\mc{X})}$---as opposed to the semi-norm $|f_0|_{H^s(\mc{X})}$--- and that $p \in C^{s - 1}(\mc{X})$. Both requirements are related and fundamental, as we shall now explain. As previously mentioned, $L$ converges to a weighted-version of the Laplace operator $\Delta_P$ (also known as a Fokker-Planck) operator. The operator $\Delta_P$ includes both first and second-order derivatives,
	\begin{equation}
	\label{eqn:fokker_planck_1}
	\Delta_Pf= \frac{-1}{p} \mathrm{div}(p \nabla f) = \Delta f - \frac{\nabla p^{\top} \nabla f}{p},
	\end{equation}
	and therefore a function $f \in H_0^s(\mc{X})$ has finite energy $\|\Delta_P f\|_{P}^2$ only if the first-order partial derivatives of $f$ and $p$ are bounded. \textcolor{red}{This discussion---and in particular the last sentence---are obviously not quite correct. This is in part because you realize now that the requirements are not in fact fundamental. Instead, the requirement should merely be that $\|\Delta_P^kf\|_{\Leb^2(P)}^2$ is bounded---here $k = s/2$ and we assume $s$ is even---and the upper bounds should be in terms of this semi-norm.}.
	\item \textbf{(1.4) Both of the remarks after Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} continue to apply.}
\end{itemize}

\textbf{(1.4) Rates for testing with higher-order smoothness assumptions.} For testing, we can get the optimal rate $n^{-4s/(4s + d)}$ when $d \leq 4$. When $4 < d \leq 4s$, we get the rate~\textcolor{red}{(...)} Recall that the rate $n^{-4s/(4s + d)}$ is optimal whenever $d \leq 4s$. Therefore there are scalings of $s$ and $d$ for which we do not get the optimal rate. We do not know whether this due to looseness in our proof techniques, or a problem with the test itself.

\subsection{Analysis}
\label{sec:analysis}

\textbf{(1) Bias and variance terms, for estimation and testing}. We analyze the estimation error of $\wh{f}$ by first condition on the design points $X$ and computing \emph{design-dependent} bias and variance terms. For estimation, we have
\begin{equation*}
\Ebb\bigl[\|\wh{f} - f_0\|_n^2|X\bigr] \leq C\frac{\dotp{L^s f_0}{f_0}_n}{\lambda_{\kappa}^s} + \frac{\kappa}{n}
\end{equation*}
For testing, we have~\textcolor{red}{(...)}.

In the above, the semi-norm $\dotp{L^s f_0}{f_0}_n$ and eigenvalue $\lambda_{\kappa}$ are random variables that depend the design $X$. It remains to derive suitable upper and lower bounds on these quantities.

\textbf{(2) Upper bound on the semi-norm. } Lemma~\ref{lem:graph_seminorm} gives a sufficient upper bound on $\dotp{L^s f_0}{f_0}_n$. Take $\beta := s - 1$. 
\begin{lemma}
	\label{lem:graph_seminorm} 
	Suppose $f \in H_0^s(\mc{X})$, that $p \in C^{\beta}(\mc{X})$, and that $r \gtrsim n^{-1/d}$. Then with probability at least $1 - \delta$, it holds that 
	\begin{equation}
	\label{eqn:graph_seminorm_1}
	\dotp{L^s f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^s} \|p\|_{C^{\beta}} + \textcolor{red}{(?)}
	\end{equation}
	Therefore, if additionally $r \gtrsim n^{-1/(2\beta + d)}$, then 
	\begin{equation}
	\label{eqn:graph_seminorm_2}
	\dotp{L^s f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^s} \|p\|_{C^{\beta}}
	\end{equation}
\end{lemma}

\textbf{(3) Lower bound on the eigenvalue.} On the other hand, recent work \textcolor{red}{(Burago14, Shi16, Garcia Trillos 18)} has analyzed the convergence of $\lambda_{k}$ towards $\lambda_{k}(\Delta_P)$. They provide explicit bounds on the relative error $|\lambda_{k} - \lambda_{k}(\Delta_P)|/\lambda_{k}(\Delta_P)$, showing that the relative error is small for sufficiently large $n$, small $\varepsilon$, and any $1 \leq k \leq n$ that is not too large relative to $n$. These results are actually stronger than are necessary to establish Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_estimation_ho}---note that we need only show that $\lambda_{\kappa}$ is lower bounded by the right order of magnitude---but unfortunately they all assume $P$ is supported on a manifold without boundary. We will instead use the results of \textcolor{red}{(Green 2021)}, whose assumptions match our own, and who give a weaker bound on $\lambda_k/\lambda_k(\Delta_P)$ that will nevertheless suffice for our purposes.

\begin{lemma}
	\textcolor{red}{(TODO)}
	With probability at least $1 - \textcolor{red}{(?)}$,
	\begin{equation*}
	\lambda_k \geq c \cdot \min\Bigl\{\lambda_k(\Delta_P), \frac{1}{r^{2}} \Bigr\}
	\end{equation*}
\end{lemma}
By Weyl's Law, $c k^{2/d} \leq \lambda_{k}(\Delta_P) \leq Ck^{2/d}$, and as a result $\lambda_{\kappa} \geq C\lambda_{\kappa}(\Delta_P)$ as long as $r \geq \kappa^{-1/d}$.

\textbf{(1.2) Traps we avoid.} It is tempting to use recent results regarding spectral convergence to analyze the estimator $\wh{f}$ and test $\wh{T}$. This suffers from an accumulation of error problem.

It is also key that we directly analyze the semi-norm $\dotp{L^s f}{f}_n$, rather than invoking the pointwise convergence of $L^{s}f \to \Delta_P^{s}f$ to obtain a bound on $\dotp{L^s f}{f}_n$. Recall that we assume $f$ has bounded derivatives up to (and including) order $s$. This suffices to gain control of the semi-norm $\dotp{L^s f}{f}_n$, because $\dotp{L^s f}{f}_n$ approximates the $\Leb^2$ norm of an order $s$ differential operator, and so we have exactly as many derivatives as we need. On the other hand, $L^s$ approximates an order $2s$ differential operator, and $\Delta_P^sf$ is only a well-defined limiting object if we assume $f$ has order $2s$ derivatives.

\section{Manifold Adaptivity}
\label{sec:manifold_adaptivity}

The optimal rates for nonparametric regression suffer from a bona fide curse of dimensionality, and are thus difficult to reconcile with the practical success of many methods for nonparametric regression. One explanation for this seeming paradox is the \emph{manifold hypothesis}, where the design points $X$ are assumed to lie on or near a manifold $\mc{X}$ of intrinsic dimension $m \ll d$. Under the manifold hypothesis nonparametric regression is demonstrably easier, and in particular it is known \citep{bickel2007,ariascastro2018} that in this setting the optimal rates over $H^s(\Xset)$ scale like $n^{-2s/(2s + m)}$ (for estimation) and $n^{-4s/(4s + m)}$ (for testing).

On the other hand, a theory has been developed~\citep{belkin03,belkin05,niyogi2013} establishing the the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear. In this section, we contribute to this line of work by showing that under the manifold hypothesis, Laplacian eigenmaps achieve the sharper minimax estimation and testing rates. First, we give some relevant background on Riemmanian manifolds, and Sobolev classes defined upon them.

\subsection{Riemmanian manifolds}

\subsection{Error rates under the manifold hypothesis}

\section{Out-of-sample error}
\label{sec:out_of_sample}
Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity} highlights the statistical optimality of nonparametric regression with Laplacian eigenmaps. However, with regards to estimation, the error has always been measured in $\|\cdot\|_n$ norm---in words, the average squared loss incurred over the $n$ observed design points. This is somewhat unusual. Instead, it is more typical to consider $L^2(P)$ error---meaning error with respect to the $\|\cdot\|_{P}$ norm---in large part because $L^2(P)$ error is closely tied to the expected prediction error an estimator would make at a new point $X_{n + 1}$.

Of course, the Laplacian eigenmaps estimator is only defined at the observed design points $X_1,\ldots,X_n$, and to measure its error in $L^2(P)$ norm we must first give a means of extending it to the rest of $\Xset$. We propose a simple method, essentially kernel smoothing, to do the job, which can be applied to any estimator at data, including Laplacian eigenmaps. We show that the smoothed version of $\wh{f}$ has optimal $L^2(P)$ error.

\subsection{Extending Laplacian Eigenmaps with Kernel Smoothing}

\begin{theorem}
	\label{thm:kernel_smoothing}
	\textcolor{red}{(TODO)}
\end{theorem}

Some remarks:
\begin{itemize}
	\item \textbf{Tuning parameters by sample-splitting.}
	\item \textbf{Other extensions.} Comment that there does not exist any canonical way to extend Laplacian based estimators such as $\wh{\theta}$ to an ambient domain. Point out that there exist various proposals to extend eigenvectors in the literature, such as $1$-nearest-neighbor regression (piecewise constant extrapolation over Voronoi cells), convolution with a specialized kernel, or Nystr\"{o}m extension. Explain that each of these is designed for a highly specialized purpose, such as suitability for mathematical analysis or computational efficiency, and in any case are suggested only for eigenvectors rather than a regression estimate. State explicitly that we choose kernel smoothing for its simplicity, ubiquity, and strong theoretical properties under our assumptions, but that considerations of extensions for specifically tailored to Laplacian based estimators could be of interest.
\end{itemize}

\subsection{Why not just do kernel smoothing?}
On the other hand, the allusion to kernel smoothing raises a very natural question: why should we prefer to first compute $\wh{\theta}$ and smooth it out, as opposed to simply performing kernel smoothing over the original responses $Y$. We now answer this question, by \textcolor{red}{(...)}.

\section{Experiments}
\label{sec:experiments}

\section{Discussion}
\label{sec:discussion}

\bibliographystyle{plainnat}
\bibliography{../../../graph_testing_bibliography}

\end{document}

### 2021-02-11 code/configs/estimation_mse_config3.R
Updated code folder.

@@ -4,7 +4,7 @@
 
 ### General configs. ###
 d <- 2
-ns <- round( 10^seq(2,3.5,length.out = 10) )
+ns <- round( 10^seq(3,4,length.out = 10) )
 iters <- 5
 
 ### Configs for sampling. ###

### 2021-02-11 code/configs/estimation_mse_config4.R

@@ -4,7 +4,7 @@
 
 ### General configs. ###
 d <- 3
-ns <- round( 10^seq(2.5,3.5,length.out = 10) )
+ns <- round( 10^seq(3,4,length.out = 10) )
 iters <- 5
 
 ### Configs for sampling. ###

### 2021-02-11 code/configs/estimation_mse_config5.R

@@ -4,7 +4,7 @@
 
 ### General configs. ###
 d <- 4
-ns <- round( 10^seq(2.5,3.5,length.out = 10) )
+ns <- round( 10^seq(3,4,length.out = 10) )
 iters <- 5
 
 ### Configs for sampling. ###

### 2021-02-11 code/configs/estimation_mse_config6.R

@@ -1,10 +1,10 @@
 #------------------------------------------------#
-# Configs for second estimation MSE experiment.
+# Configs for fifth estimation MSE experiment.
 #------------------------------------------------#
 
 ### General configs. ###
-d <- 1
-ns <- round( 10^seq(2.5,3.5,length.out = 10) )
+d <- 5
+ns <- round( 2*10^seq(2.5,3.5,length.out = 10) )
 iters <- 5
 
 ### Configs for sampling. ###

### 2021-02-11 code/configs/estimation_mse_config_linear1d.R

@@ -4,12 +4,12 @@
 
 ### General configs. ###
 d <- 1
-ns <- round( 10^seq(2.5,3.5,length.out = 10) )
+ns <- c(200,1000)
 iters <- 5
 
 ### Configs for sampling. ###
 sample_X <- make_sample_uniform(d)
-make_f0 <- make_cosine_f0
+make_f0 <- make_linear_f0
 
 
 ### Methods. ###

### 2021-02-11 code/configs/estimation_mse_config_linear2d.R

@@ -3,13 +3,13 @@
 #------------------------------------------------#
 
 ### General configs. ###
-d <- 1
+d <- 2
 ns <- round( 10^seq(2.5,3.5,length.out = 10) )
 iters <- 5
 
 ### Configs for sampling. ###
 sample_X <- make_sample_uniform(d)
-make_f0 <- make_cosine_f0
+make_f0 <- make_linear_f0
 
 
 ### Methods. ###

### 2021-02-11 code/configs/estimation_mse_config_linear3d.R

@@ -3,13 +3,13 @@
 #------------------------------------------------#
 
 ### General configs. ###
-d <- 1
+d <- 3
 ns <- round( 10^seq(2.5,3.5,length.out = 10) )
 iters <- 5
 
 ### Configs for sampling. ###
 sample_X <- make_sample_uniform(d)
-make_f0 <- make_cosine_f0
+make_f0 <- make_linear_f0
 
 
 ### Methods. ###

### 2021-02-11 code/configs/estimation_mse_config_linear4d.R

@@ -3,13 +3,13 @@
 #------------------------------------------------#
 
 ### General configs. ###
-d <- 1
+d <- 4
 ns <- round( 10^seq(2.5,3.5,length.out = 10) )
 iters <- 5
 
 ### Configs for sampling. ###
 sample_X <- make_sample_uniform(d)
-make_f0 <- make_cosine_f0
+make_f0 <- make_linear_f0
 
 
 ### Methods. ###

### 2021-02-11 notes/graph_regression.tex
Cleaned up notes folder.

--- 2020-09-16 notes/graph_regression2.tex
@@
 	Assume~\ref{asmp:kernel} and~\ref{asmp:domain}, and additionally that $p(x) \leq p_{\max}$ for all $x \in \Xset$.
-Then, for any $f \in H^1(\Xset,M)$, there exists a constant $C_1 > 0$ which depends only on $\Xset$, $d$, $p_{\max}$ and $K$ such that
+Then, for any $f \in H^1(\Xset)$, there exists a constant $C_1 > 0$ which depends only on $\Xset$, $d$, $p_{\max}$ and $K$ such that
 	\begin{equation}
@@
 \subsubsection{Testing}
-
 Let us define two simple tests, the first using $T_{\LE}$ and the second using $T_{\LS}$. For $b \geq 1$, define the thresholds $\wh{t}_b$ and $\wt{t}_b$ to be
@@
 \bibliographystyle{plainnat}
-\bibliography{../graph_testing_bibliography}
+\bibliography{../graph_regression_bibliography}
 

### 2021-02-11 notes/graph_testing.tex
Cleaned up notes folder.

@@
 \newcommand{\wt}[1]{\widetilde{#1}}
+\newcommand{\wh}[1]{\widehat{#1}}
 \newcommand{\ol}[1]{\overline{#1}}
@@
 The manner in which we do this depends on whether $\wb{x}$ is sufficiently in the interior, or conversely is near the boundary, of $\wb{X}$.
-With an appropriate approximation in hand, we then find the desired estimate on the discrete semi-norm $f^T \wb{L}^s f$.
+With an appropriate approximation of $\wb{L} f(\wb{x})$ in hand, we then find the desired estimate on the discrete semi-norm $f^T \wb{L}^s f$.
 
@@
 \subsection{Proof of Proposition~\ref{prop:L4_testing_rate}}
-
 Let $T_{\mathrm{mean}} = \frac{1}{n}\sum_{i = 1}^{n} y_i^2$. The expectation of $T_{\mathrm{mean}}$ is
@@
 
-For Lemmas~\ref{lem:expected_first_order_seminorm} - \ref{lem:expected_first_order_seminorm_3}, we will assume that $K$ is a kernel function compactly supported on $B(0,1)$ and upper bound $K(x) \leq K_{\max}$.
+For Lemmas~\ref{lem:expected_first_order_seminorm} - \ref{lem:expected_first_order_seminorm_3}, we will assume that $K$ is a kernel function compactly supported on $B(0,1)$ satisfying the upper bound $K(x) \leq K_{\max}$ for all $x \in B(0,1)$.
 Additionally, we note that although each Lemma assumes $g \in C^1(\Xset)$, in each case if we assume merely $g \in H^1(\Xset)$ the results still hold due to the density of $C^{\infty}$ in $H^1(\Xset)$.

### 2021-02-11 notes/references/reference_notes.tex
Cleaned up notes folder.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{\mathcal{L}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\Unq}{\mathrm{Unq}}
\newcommand{\rg}{\mathrm{rg}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Reference notes for the graph testing project}
\author{Alden Green}
\date{\today}
\maketitle

This document contains notes on papers that are potentially relevant to the graph testing project. I'm not sure what the organizational structure will be yet, as I haven't seriously attempted to keep a document like this before. Perhaps the top heading should be by subject matter.

\section{Convergence of Neighborhood Graph Operators}

We observe points $X = \{x_1,\ldots,x_n\}$ sampled independently from a distribution $P$ on $\Reals^d$, supported on $\Xset = [0,1]^d$. Let $K_r(x,y) = \frac{1}{r^d} K(\norm{x - y})$ be a kernel function with a given radius $r > 0$. The neighborhood graph $G_{n,r} = (V,W)$ is an undirected, weighted graph, with data points $V = X$ as vertices, and weighted edge $w_{ij} = K_r(x_i,x_j)$. The combinatorial Laplacian $L = D - W$ is an $n \times n$ matrix defined on $G_{n,r}$; here $D$ is the diagonal degree matrix with $D_{ii} = \sum_{i \neq j} W_{ij}$. The graph Laplacian operator $L_{n,r}$ is defined by the following action:
\begin{equation*}
L_{n,r}f(x) = \frac{1}{nr^2}\sum_{i = 1}^{n}(f(x_i) - f(x))K_r(x_i,x)
\end{equation*}
The nonlocal $P$-weighted Laplacian operator $L_{\Pbb,r}$ is defined by the action
\begin{equation*}
L_{\Pbb,r}f(x) = \int (f(y) - f(x))K_r(y,x) \,dP(x)
\end{equation*}

The normalized graph Laplacian operator $U_{n,r}$ is defined as
\begin{equation*}
U_{n,r} := I - T_{n,r},~~ T_{n,r}f := \int h_{n,r}(x,y)f(y)\,dP_n(y), ~~h_r(x,y) := \frac{K_r(x,y)}{\sqrt{d_n(x)d_n(y)}}
\end{equation*}
and the normalized $P$-weighted Laplacian operator $U_{P,r}$ is defined as
\begin{equation*}
U_{P,r} := I - T_{P,r},~~ T_{P,r}f := \int h_{P,r}(x,y)f(y)\,dP_n(y), ~~h_r(x,y) := \frac{K_r(x,y)}{\sqrt{d(x)d(y)}}
\end{equation*}

We denote the spectrum of any operator (matrix) $T$ by $\lambda(T)$. When $T$ is a matrix, we let $\lambda_k(T)$ denote the $k$th smallest eigenvalue of $T$.

\subsection{Consistency of Spectral Clustering (vonLuxburg, Belkin, Bousquet 04)}

\textit{Relevance to our work:} 
This paper establishes, under appropriate conditions, the convergence $\lambda_k(L) \to \lambda_k(L_{P,r})$ for a fixed $k$ and fixed graph radius $r$. It also establishes the convergence of eigenvectors in $L^{\infty}$. For application to supervised learning problems where we intend to take a growing number of eigenvalues, we would like to show that both hold for some growing value of $k$.

\textit{Summary:} We seek to examine the proof of the following Theorem, regarding the convergence of the eigenvalues $\lambda_k(L)$. Let $d: \Xset \to \Reals$ be the degree function,
\begin{equation*}
d(x) = \int K_r(x,y) \,dP(y),~~ d_n(x) = \int K_r(x,y) \,dP_n(y)
\end{equation*}
and write $\rg(f)$ for the range of a function $f$.

\begin{theorem}
	\label{thm:vonluxburg04_1}
	Let $\lambda \neq \rg(d)$ be an eigenvalue of $L_{P,r}$ and $M \subset \mathbb{C}$ be an open neighborhood of $\lambda$ such that $\lambda(U) \cap \bar{M} = \set{\lambda}$, Then every sequence $(\lambda_n)$ with $\lambda_n \in \lambda(\frac{1}{nr^2}L) \cap M$ satisfies $\lambda_n \to \lambda$ almost surely.
\end{theorem}

The following theorem establishes that the rate of convergence of the eigenvectors of the normalized graph Laplacian is no slower than the supremum of a particular empirical process. The function class will depend on
\begin{equation*}
\mathcal{K} := \Bigl\{ K_r(x,\cdot): x \in \Xset \Bigr\}, ~~ \mathcal{H} := \Bigl\{h(x,\cdot): x \in \Xset\Bigr\}.
\end{equation*}
\begin{theorem}
	\label{thm:vonluxburg04_2}
	Under the gneral assumptions, let $\lambda \neq 0$ be a simple eigenvalue of $T$, with corresponding eigenfunction $u$. Let $(\lambda_n)$ be a sequence of eigenvalues of $(T_n)$ which satisfy $\lambda_n \to \lambda$, with corresponding eigenvectors $(u_n)$. Then, for $\mathcal{F} = \mathcal{K} \cup u \cdot \mathcal{ H} \cup \mathcal{H} \cdot \mathcal{H}$, there exists a constant $C$, which depends only on $K$, $r$, $\lambda$ and $\sigma(T)$, and a sequence of signs $(a_n) \subset \{+1,-1\}$, such that
	\begin{equation*}
	\norm{a_nu_n - u}_{\infty} \leq C \sup_{f \in \mathcal{F}} \abs{P_n(f) - P(f)}
	\end{equation*}
\end{theorem}

\subsubsection{Proofs}

\paragraph{Proof of Theorem~\ref{thm:vonluxburg04_1}:}

We will show that $L_{n,r} \overset{c}{\to} L_{P,r}$. This implies Theorem~\ref{thm:vonluxburg04_1} by the following argument. Suppose $(\lambda_n)$ is a sequence of eigenvalues of $L$, such that $\lambda_n \in \bar{M}$ for every sufficiently large $n$. Since $\lambda_n \in \bar{M}$, $\lambda_n \not\in \rg(d)$. Therefore for sufficiently large $n$ $\lambda_n \not\in \rg(d_n)$, since $d_n$ converges in operator norm to $d$ and so $\rg(d_n) \to \rg(d)$. This implies $\lambda_n$ is an eigenvalue of $L_{n,r}$ for all sufficiently large $n$, as the eigenvalues of $L$ and $L_{n,r}$ are one-to-one outside of the essential spectrum of $L_{n,r}$. The Theorem then follows by reasoning along the lines of Section~\ref{sec:modes_of_convergence}.

It remains to show $L_{n,r} \overset{c}{\to} L_{P,r}$. Let 
\begin{equation*}
M_{d_n}f(x) = f(x) d_n(x),~~T_nf(x) = \int K_r(x,y) f(y) \,dP_n(y)
\end{equation*}
and similarly define $M_{d_P}$ and $T_P$ with respect to the underlying measure $P$. Since $L_{n,r} = M_{d_n} - T_n$ and likewise $L_{P,r} = M_{d_P}- T_{P}$, it will suffice to show that (i) $M_{d_n} \overset{\mathrm{op}}{\to} M_{d_P}$ and (ii) $T_n \overset{\mathrm{c}}{\to} T_P$.

We show operator norm convergence of $M_{d_n}$ to $M_{d_P}$ by reducing $\norm{M_{d_n} - M_{d_P}}_{\mathrm{op}}$ to an empirical process term. Letting $\mathcal{K} = \{K(x,\cdot): x \in \Xset\}$, we have
\begin{align}
\norm{M_{d_n} - M_{d_P}}_{\mathrm{op}} & = \sup_{f \in B} \norm{M_{d_n}f - M_{d_P}f}_{\infty} \\
& = \sup_{f \in B} \sup_{x \in \Xset}\bigl\{M_{d_n}f(x) - M_{d_P}f(x)\bigr\} \\
& \leq \sup_{x \in \Xset}\bigl\{d_n(x) - d(x)\bigr\} \\
& = \sup_{K_x \in \mathcal{K}} \abs{P_n(K_x) - P(K_x)} 
\end{align}
Then since $K$ is continuous, and $\Xset$ is compact, $K$ is uniformly continuous. As such, for any $\epsilon > 0$ one can construct a finite $\epsilon$-cover of $\mathcal{K}$ out of a $\delta$-cover of $\Xset$ for suitably small $\delta$. This implies $\mathcal{K}$ has a finite $\norm{\cdot}_{L_1(P)}$ covering numbers, and therefore it is Glivenko-Cantelli.

Now we show compact convergence of $T_n$ to $T_P$. We begin by showing that $T_n$ converges pointwise to $T_P$ over $C(\Xset)$. This follows since
\begin{equation*}
\abs{T_nf - T_Pf}_{\infty} = \sup_{g \in f \cdot \mathcal{K}} \abs{P_n(g) - P(g)}
\end{equation*}
where $f \cdot \mathcal{K} = \set{f \cdot K_x: x \in \Xset}$ is Glivenko-Cantelli by a similar argument to the preceding paragraph.

Finally, we show that $\cup T_nB$ is precompact, which implies that $T_nx_n$ is precompact for any $(x_n) \subset B$. Fix $x \in X$. First, we have that
\begin{equation*}
\abs{T_nf(x)} \leq \norm{K(x,\cdot)}_{\infty}
\end{equation*}
which is upper bounded by assumption. Second, we have that
\begin{equation*}
\abs{T_nf(x) - T_nf(y)} = \abs{\int f(z)\bigl(K_r(x,z) - K_r(y,z)\bigr)\,dP_n(z)} \leq \norm{K_r(x,\cdot) - K_r(y,\cdot)}_{\infty}
\end{equation*}
and the second is upper bounded by the uniform continuity of $K_r$. The relative compactness of $\cup T_nB$ follows from Arzela-Ascoli.

\paragraph{Proof of Theorem~\ref{thm:vonluxburg04_2}:}
We start with a result from \textcolor{red}{Atkinson}.
\begin{theorem}
	\label{thm:atkinson_spectral_perturbation}
	Let $(E,\norm{\cdot}_E)$ be an arbitrary Banach space and $B$ its unit ball. Let $(K_n)$ and $K$ be compact linear operators such that $K_n \overset{cc}{\to} K$. For a nonzero eigenvalue $\lambda \in \sigma(K)$, denote the corresponding spectral projection by $\mathrm{Pr}$. Let $M \subset \mathbb{C}$ be an open neighborhood of $\lambda$ such that $\sigma(K) \cap M = \set{\lambda}$. Then there exists a constant $C > 0$ such that for every $x \in \mathrm{Pr}E$, 
	\begin{equation*}
	\norm{x - \mathrm{Pr}_nx}_{E} \leq C\Bigl( \norm{(K_n - K)x}_{E} + \norm{x}_E \cdot \norm{(K - K_n)K_n} \Bigr)
	\end{equation*}
	where $C$ is independent of $x$, but depends on $\lambda$ and $\sigma(K)$.
\end{theorem}
To apply the result of Atkinson, we make use of the following proposition.
\begin{proposition}
	Let $v \in E$ be a unit norm vector. Let $(v_n)$ with $v_n \in E$ be a sequence of vectors with $\norm{v_n} = 1$, and let $\Pr_n$ be the corresponding spectral projections. Then there exists some $(a_n) \subset \{+1,-1\}$ such that
	\begin{equation*}
	\norm{a_nv_n - v}_E \leq 2\norm{\Pr_nv - v}
	\end{equation*}
	for each $n$.
\end{proposition}
\begin{proof}
	By the triangle inequality,
	\begin{equation*}
	\norm{a_nv_n - v} \leq \norm{a_nv_n - \Pr_nv} + \norm{\Pr_nv - v}
	\end{equation*}
	Then, there exists $a_n \in \{+1,-1\}$ such that
	\begin{align*}
	\norm{a_nv_n - \Pr_nv} & = 1 - \norm{\Pr_nv} \\
	& = 1 - \norm{v - (v - \Pr_nv)} \\
	& \leq 1 - \abs{\norm{v} - \norm{v - \Pr_nv}} \\
	& = \norm{v - \Pr_nv}
	\end{align*}
	and by the reverse triangle inequality $\norm{\Pr_nv} = \norm{v - (v - \Pr_nv)} \geq \abs{\norm{v} - \norm{v - \Pr_nv}} = 1 - \norm{v - \Pr_nv}$. 
\end{proof}

\section{Modes of Convergence}
\label{sec:modes_of_convergence}

Where I play with mathematical notions currently beyond my limits.

Let $(T_n)$ and $T$ be closed operators on a Banach space $(E,\norm{\cdot}_E)$. Denote the unit ball of $E$ by $B$. We say 
\begin{itemize}
	\item $T_n \overset{p}{\to} T$ if $\norm{(T_n - T)x}_E \to 0$ for every $x \in E$. 
	\item $T_n \overset{c}{\to} T$ if (i) $T_n \overset{p}{\to} T$ and (ii) for every $(x_n) \subset B$, the sequence $(T - T_n)x_n$ is precompact.
	\item $T_n \overset{c}{\to} T$ if (i) $T_n \overset{p}{\to} T$ and (ii) the set $\bigcup_{n} (T_n - T)B$ is precompact.
	\item $T_n \overset{r}{\to} T$ if (i) $T_n \overset{p}{\to} T$ and (ii) for every $(x_n) \subset B$ such that $T_nx_{n} \to y$, the sequence $(x_{n})$ has a limit point $x$.
	\item $T_n \overset{s}{\to} T$ if (i) $T_n \overset{p}{\to} T$ and (ii) there exists $M > 0, N$ such that for all $n \geq N$, $\norm{T_n^{-1}} \leq M$. 
\end{itemize}

\textbf{Claim 1: $T_n \overset{c}{\to} T$ implies $T_n - z \overset{r}{\to} T - z$ for any $z$ in the resolvent of $T$.}

Suppose $(x_n) \subset B$ is a sequence satisfying $(T_n - z)x_n \to y$. Then, $(T - z)x_n = (T - T_n)x_n + (T_n - z)x_n$. The first converges along a subsequence to some $u$ by hypothesis. The second converges to some $y$ by the previous assumption. Therefore, $(T - z)x_n$ converges along a subsequence to $y + u$. Since $z$ is in the resolvent of $T$, this implies $(T - z)^{-1}$ exists, and therefore $x_n$ converges along a subsequence to $x = (T - z)^{-1}(y + u)$. Then $(T - T_n)x_n = (T - T_n)x + (T-T_n)(x - x_n)$, with the former converging to $0$ by the pointwise convergence of $T_n$ to $T$, and the latter converging to $0$ along a subsequence as $T$ and $T_n$ are bounded. Therefore $u = 0$, and $(T - z)x = y$.

\textbf{Claim 2: $T_n - z \overset{r}{\to} T - z$ implies $T_n - z \overset{s}{\to} T - z$ for all $z$ in the resolvent of $T$.}

Fix $z \in \rho(T)$. It suffices to show that there exists $M > 0, N$ such that for all $(x_n)$, $\norm{(T_n - z)x_n}_{E} \geq M^{-1} \norm{x_n}_E$. Suppose this were not true. Then there would exist a sequence $(x_n), \norm{x_n}_E = 1$ such that $\norm{(T_n - z)x_n}_{E} \to 0$. Then by assumption, there exists some limit point $x$ of $(x_n)$ such that $(T - z)x = 0$. However since $z \in \rho(T)$, this implies $x = 0$. As $0$ cannot be the limit point of $(x_n)$, this is a contradiction, proving the claim.

\textbf{Claim 3: Suppose $\lambda$ is an eigenvalue of $T$ such that $\sigma(T) \cap M = \set{\lambda}$ for some $M \subset \mathbb{C}$. Then, $T_n - z \overset{s}{\to} T - z$ for all $z \in M - \set{\lambda}$ implies that any $(\lambda_n) \in \sigma(T_n) \cap M$ satisfies $\lim \lambda_n = \lambda$.}

Suppose otherwise; then there exists $(\lambda_n') \in \sigma(T_n) \cap M$ which converges along a subsequence to some $\lambda' \in \sigma(T) \cap M, \lambda' \neq \lambda$. Without loss of generality assume $\lim \lambda_n' = \lambda'$. Therefore there exists $(x_n'), \norm{x_n'}_E = 1$ such that
\begin{equation*}
\lim_{n \to \infty} (T_n - \lambda')x_n' = 0
\end{equation*}
which by regular convergence implies $x_n'$ has some limit point $x$, such that $(T - \lambda')x = 0$. Then since $\lambda'$ is in the resolvent of $T$, this implies $x = 0$, which is a contradiction.

Since $T_n \to T$, we also have that if $Tx = \lambda x$, $T_nx \to \lambda x$.

\section{Continuum and Discrete Norms}
This section is devoted to a better understanding of couplings between continuum and discrete norms. For example, under what conditions does a continuum Sobolev space embed nicely into a discrete Sobolev space?

\subsection{Asymptotic Optimality of the $C_p$-Test \textcolor{red}{(Polyak and Tsybakov 90)}}

\textit{Relevance to our work:} We would like to show that the discretization error due to using sums instead of integrals to compute Fourier coefficients is negligible relative to the bias-variance tradeoff, for as wide a class of functions $f$ as possible.

Let $x_i = i/n$ for $i = 1,\ldots,n$, and suppose we observe responses according to the fixed design regression model
\begin{equation*}
y_i = f(x_i) + \xi_i,~~ \Ebb(\xi_i) = 0, \Ebb(\xi_i\xi_j) = \delta_{ij} \sigma^2.
\end{equation*}
where we assume that $f$ can be represented as a series converging in $L_2[0,1]$,
\begin{equation}
f(x) = \sum_{j = 1}^{\infty} c_j \varphi_j(x),~~ x \in [0,1].
\end{equation}
We take $f_{nN}$ to be an orthogonal series estimator of $f$,
\begin{equation*}
f_{nN}(x) = \sum_{j = 1}^{N} \wh{c}_{jn} \varphi_j(x)
\end{equation*}
where $\psi_j$ is the orthonormal trigonometric system on $[0,1]$, given by
\begin{align*}
\varphi_1(x) & \equiv 1 \\
\varphi_k(x) & = \sqrt{2}\cos(2\pi k x),~~k \in \mathbb{N} \\
\varphi_k(x) & = \sqrt{2}\sin(2\pi k x),~~k \in \mathbb{N}
\end{align*}
and $\wh{c}_{jn} = \frac{1}{n}\sum_{i = 1}^{n} f(x_i) \varphi_j(x_i)$.

Let $R_{nN}^{(3)}$ be the risk of estimator $f_nN$, given by
\begin{equation*}
R_{nN}^{(3)} = \frac{1}{n} \mathbb{E}\Bigl[\sum_{i = 1}^{n} (f_nN(x_i) - f(x_i))^2\Bigr].
\end{equation*}
(We will not concern ourselves with the other notions of risk laid out in this paper.)
Suppose additionally that $c_j \in \ell^1(\mathbb{N})$, i.e. that there exists a sequence of non-negative numbers $\varepsilon_j$ such that $j\varepsilon_j$ is monotone non-increasing in $j$ and 
\begin{equation}
\label{asmp:polyak_1}
\abs{c_j} \leq \varepsilon_j,~~ \sum_{j = 1}^{\infty} \varepsilon_j = C < \infty.
\end{equation}
(Actually, this is just a tad stronger than $c_j \in \ell^1(\mathbb{N})$.)
\begin{theorem}[Theorem 1, (1.3)]
	Let $A_nN = \rho_N + \sigma^2 N / n$, where $\rho_N = \sum_{j = N + 1}^{\infty} c_j^2$. If~\eqref{asmp:polyak_1} holds, then,
	\begin{equation*}
	\frac{\max_{N \leq n}\abs{R_{nN}^{(3)} - A_{nN}}}{\min_{N}A_{nN}} \to 0,~~ \textrm{as $n \to \infty$.}
	\end{equation*}
\end{theorem}

Actually the only part I care about are the following Lemmas.
\begin{lemma}[Lemma 1 \textcolor{red}{(Polyak and Tsybakov)}]
	For $n \geq 1$, 
	\begin{equation*}
	\frac{1}{n} \sum_{m = 1}^{n} \varphi_l(x_m) \varphi_j(x_m) = \delta_{lj} + \Delta_{lj}
	\end{equation*}
	where $\Delta_{lj} = 0$ if $l,j \leq n$ or if $l$ and $j$ are of different parity, and otherwise
	\begin{equation*}
	\abs{\Delta_{lj}} \leq \sum_{k = 1}^{\infty} \Bigl(\delta_{\abs{l_1 - j_1} = kn} + \delta_{l_1 + j_1 = kn}\Bigr)
	\end{equation*}
	where $l_1 = [l/2], j_1 = [j/2]$.
\end{lemma}

\begin{lemma}[Lemma 2 \textcolor{red}{Polyak and Tsybakov}]
	\label{lem:polyak90_2}
	Suppose condition~\eqref{asmp:polyak_1} is satisfied. Then there exists a constant $K_1$ depending only on $C$ such that
	\begin{equation*}
	\max_{1 \leq j \leq n} \biggl(\sum_{k = n + 1}^{\infty} c_k \Delta_{jk}\biggr)^2 \leq \frac{K_1}{n^2}.
	\end{equation*}
\end{lemma}
\subsubsection{Proofs}
\paragraph{Proof of Lemma~\ref{lem:polyak90_2}.}
Let $j$ and $n$ be even. Then,
\begin{align*}
\abs{\sum_{k = n + 1} c_k \varDelta_k} & = \abs{\sum_{k = n + 2, k~\textrm{even}} c_k \varDelta_k} \\
& \leq \sum_{k_1 = n/2 + 1} \abs{c_{2k_1}} \sum_{m = 1}^{\infty} \Bigl(\delta_{\abs{k_1 - j_1} = mn} + \delta_{k_1 + j_1 = mn}\Bigr) \\
& = \sum_{m = 1}^{\infty} \abs{c_{2j_1 + 2mn}} + \abs{c_{2mn - 2j_1}} \\
& \leq \sum_{m = 1}^{\infty} \varepsilon_{2j_1 + 2mn} + \varepsilon_{2mn - 2j_1}
\end{align*}
Since $j_1 > n/2$, we have $2j_1 + 2mn,2mn - 2j_1 \geq m n$ for each $m \in \mathbb{N}$. Therefore by assumption, $\varepsilon_{2j_1 + 2mn}, \varepsilon_{2mn - 2j_1} \leq \varepsilon_{mn} \leq (1/n) \varepsilon_{m}$. Along with condition~\eqref{asmp:polyak_1}, this implies Lemma~\ref{lem:polyak90_2}.

\section{Random Design Regression}
In the random design regression model, we observe
\begin{equation*}
Y_i = f_0(X_i) + \varepsilon_i
\end{equation*}
where $X_i \sim P_i$ take values in space $\Xset$, $f_0$ is an unknown map from $\Xset \to \Reals$, and $\varepsilon_i$ come from a common error distribution $N$.

This section is devoted to a review of results regarding estimation risk in the random design setting. The emphasis is on papers which place as few restrictions on the distributions $P_i$ as possible.

\subsection{A New Method For Estimation and Model Selection: $\rho$-Estimation \textcolor{red}{(Baraud, Birge, and Sart 16)} }

\subsection{}
\clearpage

\end{document}

### 2021-02-11 notes/weekly_notes/11.26.20.NotesForWeek.tex
Cleaned up notes folder.

\documentclass{article}
\usepackage{amsmath, amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{verbatim}
\usepackage{fullpage}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}

% Make a widecheck symbol (thanks, Stack Exchange!)
\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{
	<5> <6> <7> <8> <9> <10>
	<10.95> <12> <14.4> <17.28> <20.74> <24.88>
	mathx10
}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareFontSubstitution{U}{mathx}{m}{n}
\DeclareMathAccent{\widecheck}{0}{mathx}{"71}
% widecheck made

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\iid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}
\newcommand{\lambdavec}{\boldsymbol{\lambda}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}
\newcommand{\Lap}{L}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}
\newcommand{\Id}{I}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Vset}{\mathcal{V}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\SM}{\mathrm{SM}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Theorem environments
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Report on: Extension of (graph Laplacian) estimators}
\author{Alden Green}
\date{\today}
\maketitle

\section{Extension of smooth functions}

We review some classical results regarding extension of smooth functions. Suppose we observe data\\ $(X_1,f^{\ast}(X_1)), \ldots, (X_n,f^{\ast}(X_n))$, where $X_1,\ldots,X_n$ are independently sampled from a distribution $P$ with support $\Omega \subset \Rd$, and $f^{\ast}: \Omega \to \Reals$ is a function smooth in some sense. Our goal is to obtain an estimate $\wt{f}: \Omega \to \Reals$, which agrees with $f^{\ast}$ at the sample points $X_1,\ldots,X_n$, and is close to $f^{\ast}$ on the rest of $\Omega$. Naturally, the size of $\wt{f} - f^{\ast}$ will depend on the \emph{fill distance} $h(X_n) := \max_{x \in \Omega} \min_{i = 1,\ldots,n} \|x - X_i\|_2$

\paragraph{Bounded variation functions, univariate.} Let $d = 1$ and $\Omega = [0,1]$. Lemma~15 in Ryan's paper on (discrete) splines establishes the error with which a function $f^{\star} \in BV([0,1])$ can be approximating using splines, assuming we observe only $(X_1,f^{\ast}(X_1)), \ldots, (X_n,f^{\ast}(X_n))$. For $k = 0$ or $k = 1$, we can use piecewise constant or piecewise linear interpolation.

\begin{lemma}
	Let $f$ be a function that is $k$ times weakly differentiable on $[0,1]$, such that $D^k f$ is of bounded variation. Also let $0 \leq x_1 < \cdots < x_n \leq 1$ be arbitrary design points. Then there exists a $k$th degree spline $g \in \mc{G}_n^k$, with knots in $x_{(k + 1):(n - 1)}$, such that for $k = 0$ or $k = 1$,
	\begin{equation*}
	\mathrm{TV}(D^k g) \leq \mathrm{TV}(D^k f),~~\textrm{and}~~ g(x_i) = f(x_i), i = 1,\ldots,n
	\end{equation*}
	and for $k \geq 2$,
	\begin{equation*}
	\mathrm{TV}(D^kg) \leq a_k \mathrm{TV}(D^k f),~~\textrm{and}~~ \|f - g\|_{\Leb_{\infty}} \leq b_k \delta_n^k \cdot \mathrm{TV}(D^k f).
	\end{equation*}
\end{lemma}
In the lemma $\mc{G}_n^k$ is the space of $k$th degree splines with knots at $x_1,\ldots,x_n$, and $x_{(k + 1):(n - 1)} = \{x_{k + 1},\ldots, x_{n - 1}\}$.

\paragraph{Sobolev functions, multivariate.}
When $f^{\ast}$ belongs to the Sobolev space $H^m(\Omega)$ for some $m > d/2$, one such method for obtaining $\wt{f}$ is \emph{thin-plate spline interpolation}. Let $If$ be the solution to
\begin{equation*}
\min |v|_{H^m(\Rd)},~~\textrm{s.t. $v(X_i) = f(X_i)$}
\end{equation*}
Then the size of the reconstruction error $I(f) - f$ can be upper bounded as a function of the \emph{fill distance} $h(X_n) := \max_{x \in \Omega} \min_{i = 1,\ldots,n} \|x - X_i\|_2$.

\begin{proposition}[Proposition 3 of J. Duchon]
	\label{prop:duchon_thin_plate_spline_approximation}
	Suppose $f \in H^m(\Omega)$. There exist constants $h_0, C_1, C_2$ depending only on $\Omega, m$, and $d$ such that for any collection $x_n = \{x_1,\ldots,x_N\} \subset \Omega$ with fill distance $h:= h(x_n) \leq h_0$, and any $k,p$ such that $W^{k,p}(\Omega) \subset H^m(\Omega)$, 
	\begin{equation*}
	\sum_{\abs{\alpha} = k} \|D^{\alpha}(f - If)\|_{L^p(\Omega)} \leq C_1 h^{m - k - d/2 + d/p} |f|_{H^m(\Omega)}
	\end{equation*}
\end{proposition}
By taking $p = \infty$, such a statement establishes a kind of \emph{local} approximation property of splines (as Ryan said). On the other hand by taking $k = 0$ and $p = 2$, we get
\begin{equation*}
\|f - If\|_{L^2(\Omega)} \leq C_1 h^{m} |f|_{H^m(\Omega)}.
\end{equation*}
If the distribution $P$ has a density upper and lower bounded away from $0$ and $\infty$, then the fill distance $h(X_n) \sim (\log n/n)^{1/d}$, and 
\begin{equation*}
\|f^{\ast} - If^{\ast}\|_{L^2(P)}^2 \leq C_2 \biggl(\frac{\log n}{n}\biggr)^{2m/d}
\end{equation*}
for a constant $C_2$ which depends only on $\Omega,m, d$ and $P$.

\section{Extension of functions in the semi-supervised setting}

We are interested in the following: if I have access only to some $g$ which approximates $f$ at the sample points $X_1,\ldots,X_n$, can I recover something akin to Proposition~\ref{prop:duchon_thin_plate_spline_approximation}? One imagines this may be known, but nevertheless I will try to prove such a fact myself. Because I am more facile with spectral graph theory than anything else, I will operate in the semi-supervised setting.

\paragraph{The semi-supervised setting.}

Suppose we observe labeled data $(X_1,\wh{\theta}_1),\ldots,(X_n,\wh{\theta}_n)$ and unlabeled data $X_{n + 1},\ldots,X_{n + m}$. Let $\wt{\theta}$ be the solution to
\begin{equation*}
\min~~\theta^{\top} L \theta,~~\textrm{s.t.}~~\theta_i = \wh{\theta}_i~~\textrm{for $i = 1,\ldots,n$.}
\end{equation*}
Here $L$ is the graph Laplacian matrix over all $m + n$ samples, meaning that for $u \in \Reals^{m + n}$,
\begin{equation*}
(Lu)_i = \sum_{j = 1}^{m + n} (u_i - u_j) K_{ij}
\end{equation*}
where $K_{ij} = K(\|X_i - X_j\|/r)$ is the $(i,j)$-th entry in the kernel matrix $K \in \Reals^{(n + m) \times (n + m)}$.  Writing 
$L =
\begin{pmatrix}
L_{n,n} & L_{n,m} \\
L_{m,n} & L_{m,m}
\end{pmatrix}
$ in block matrix notation, the estimate $\wt{\theta}$ is given by
\begin{equation*}
\wt{\theta}_i = 
\begin{cases}
\bigl(\underbrace{L_{m,m}^{-1} L_{m,n}}_{=: H} \wh{\theta}\bigr)_{i - n},~~ & i = n + 1,\ldots,n + m\\
\wh{\theta}_i,~~ & i = 1,\ldots,n
\end{cases}
\end{equation*}

\subsection{Analysis}
We would like to upper bound $\|\wh{\theta} - \theta^{\ast}\|_{m}^2$. Let $\theta_{1:n}^{\ast} = (\theta_1^{\ast},\ldots,\theta_n^{\ast})$ be the restriction of $\theta^{\ast}$ to the labeled sample points. Using the triangle inequality, we have
\begin{equation}
\|\wt{\theta} - \theta^{\ast}\|_{m}^2 \leq 2\Bigl(\|H\theta_n^{\ast} - \theta^{\ast}\|_m^2 + \|H(\wh{\theta} - \theta_n^{\ast})\|_{m}^2\Bigr).
\end{equation}
The first term here is the interpolation error, and (it would seem) should be amenable to similar analysis and bounds as Proposition~\ref{prop:duchon_thin_plate_spline_approximation}. The second term here is like a blurred version of the estimation error $\|\wh{\theta} - \theta^{\ast}\|_n^2$, and in Lemma~ we show that this blurring does not substantially increase the error, under the following conditions on the kernel $K$:
\begin{enumerate}[label=(A\arabic*)]
	\item 
	\label{asmp:kernel} The kernel $K: [0,\infty) \to \Reals$ has support $\mathrm{supp}(K) \subseteq [0,1]$, and $\sup_{x \in [0,1]} K(x) \leq C_K < \infty$.
\end{enumerate}
The Lemma will be stated with respect to the following functionals:
\begin{align*}
d_{\max}^{(n)} & := \max_{j = n + 1,\ldots,m + n} \sum_{i = 1}^{n} K_{ij},~~ && \wb{d}_{\max}^{(n)} := \max_{j = n + 1,\ldots,m + n} \sum_{i = 1}^{n} 1_{[0,1]}\biggl(\frac{\|X_i - X_j\|}{2r}\biggr), \\
d_{\max}^{(m)} & := \max_{i = 1,\ldots,n} \sum_{j = n + 1}^{m + n} K_{ij},~~ && d_{\min}^{(n)} := \min_{j = n + 1,\ldots,m + n} \sum_{i = 1}^{n} K_{ij}
\end{align*}
\begin{lemma}
	\label{lem:blurred_error}
	Under assumption~\ref{asmp:kernel}, it holds that
	\begin{equation}
	\label{eqn:blurred_error}
	\|H(\wh{\theta} - \theta^{\ast})\|_m^2 \leq C_K \frac{n}{m}\frac{d_{\max}^{(m)} \wb{d}_{\max}^{(n)}}{(d_{\min}^{(n)})^2}\|\wh{\theta} - \theta_{1:n}^{\ast}\|_n^2.
	\end{equation}
\end{lemma}

Note that under standard conditions on $P$ and $r$, $d_{\max}^{(m)} = \Theta_P(mr^d)$, and $d_{\min}^{(n)}, \wb{d}_{\max}^{(n)} = \Theta_P(nr^d)$. Under such conditions~\eqref{eqn:blurred_error} implies
\begin{equation*}
\|H(\wh{\theta} - \theta^{\ast})\|_m^2 = O_P(1) \cdot \|\wh{\theta} - \theta_{1:n}^{\ast}\|_n^2
\end{equation*}

\begin{proof}[Proof of Lemma~\ref{lem:blurred_error}]
	The proof will be completely independent of the properties of $\wh{\theta}$ or $\theta_{1:n}^{\ast}$. Letting $e := \wh{\theta} - \theta_{1:n}^{\ast}$, we have
	\begin{equation*}
	\|He\|_m^2 = \frac{1}{m} e^{\top} L_{n,m} L_{m,m}^{-2} L_{m,n} e.
	\end{equation*}
	Note that $L_{m,m} = L^{(m)} + D$, where $L^{(m)}$ is the graph Laplacian matrix w.r.t to the graph $G^{(m)}$, and $D \in \Reals^{m \times m}$ is a diagonal matrix with entries $D_{jj} = \sum_{i = 1}^{n}K_{i,(j + n)}$. Since $L^{(m)}$ is positive definite and $D \succeq d_{\min}^{(n)} I$, it holds that $L_{m,m} \succcurlyeq d_{\min}^{(n)} I$, or equivalently $L_{m,m}^{-1} \preccurlyeq \frac{1}{d_{\min}^{(n)}} I$. Hence,
	\begin{equation*}
	\frac{1}{m} e^{\top} L_{n,m} L_{m,m}^{-2} L_{m,n} e \leq \frac{1}{m}\|L_{m,n}e\|_{2}^2 \biggl(\frac{1}{d_{\min}^{(n)}}\biggr)^2.
	\end{equation*}
	Now, we can rewrite $\|L_{m,n}e\|_{2}^2$ as a quadratic form:
	\begin{equation*}
	\|L_{m,n}e\|_{2}^2 = \sum_{i = 1}^{n} \sum_{\ell = 1}^{n} e_i e_{\ell} \underbrace{\biggl\{\sum_{j = n + 1}^{n + m} K_{i,j + n} K_{\ell,j + n}\biggr\}}_{=: \eta_r^{(m)}(X_i,X_{\ell})},
	\end{equation*}
	and letting $\mc{E}$ be the matrix with entries $\mc{E}_{ij} = \eta_r^{(m)}(X_i,X_{\ell})$, we observe that
	\begin{equation*}
	\|L_{m,n}e\|_{2}^2 \leq \|e\|_2^2 \cdot \lambda_{\max}(\mc{E}) \leq \|e\|_2^2 \cdot \max_{i = 1,\ldots,n} \Bigl\{\sum_{\ell = 1}^{n} \mc{E}_{i\ell}\Bigr\}
	\end{equation*}
	where the latter inequality follows from the Perron-Frobenius inequality. Now, invoking our assumptions on the kernel function $K$, we see that $K_{i,j + n} K_{\ell,j + n} \leq C_K K_{i,j + n}$, and moreover by the triangle inequality $K_{i,j + n} K_{\ell,j + n} = 0$ unless $\|i - \ell\| \leq 2r$. Therefore, for any $i = 1,\ldots,n$, we deduce that
	\begin{equation*}
	\sum_{\ell = 1}^{n} \mc{E}_{i\ell} \leq C_k \biggl\{\sum_{\ell = 1}^{n} \1_{[0,1]}\biggl(\frac{\|X_i - X_\ell\|}{2r}\biggr) \biggr\} \cdot \max_{i = 1,\ldots,n} \biggl\{\sum_{j = n + 1}^{n + m} K_{i,j + n}\biggr\}
	\end{equation*}
	completing the proof.
\end{proof}
\end{document}

### 2021-02-11 notes/weekly_notes/8.26.20.NotesForWeek.tex
Cleaned up notes folder.

@@
 This will be the easiest of the three Lemmas to prove, while still giving some sense of how the proofs develop in the more complicated cases.
-The upper bound will depend on the kernel $K$; we assume throughout that $K:[0,\infty) \to [0,\infty)$ is a $K_{\max}$-Lipschitz function supported on $[0,1]$, and without loss of generality that
+The upper bound will depend on the kernel $K$ through the quantity
 \begin{equation*}
-\int_{\Reals^d} K(\norm{z}) \,dz = 1.
+\sigma_K := \frac{1}{d}\int_{\Reals^d} \|z\|^2 K\bigl(\|z\|\bigr) \,dz
 \end{equation*}
+
 \begin{lemma}
 	\label{lem:first_order_graph_sobolev_seminorm}
-	Suppose $\Xset$ is a Lipschitz domain, and that $p(x) \leq p_{\max}$ for all $x \in \Xset$.
-Then, for any $f \in H^1(\Xset,M)$, there exists a constant $C > 0$ which may depend only on $\Xset$ and $d$ such that
+	For any $f \in H^1(\Xset)$, we have that
 	\begin{equation}
-	f^T \Lap_{n,r} f \leq \frac{C}{\delta} n^2 r^{d + 2} M^2
+	f^T \Lap_{n,r} f \leq \frac{p_{\max}^2 \sigma_K}{\delta} n^2 r^{d + 2} |f|_{H^1(\Xset)}^2
 	\end{equation}
@@
 	\begin{equation*}
-	\Ebb\Bigl[f^T \Lap_{n,r} f\Bigr] \leq C n^2 r^{d + 2} M^2,
+	\Ebb\Bigl[f^T \Lap_{n,r} f\Bigr] \leq p_{\max}^2 \sigma_K n^2 r^{d + 2} |f|_{H^1(\Xset)}^2
 	\end{equation*}
@@
 	\begin{equation*}
-	f^T \Lap_{n,r} f = \frac{1}{2}\sum_{i, j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 A_{ij},
+	f^T \Lap_{n,r} f = \frac{1}{2}\sum_{i, j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 \mathbf{W}_{ij},
 	\end{equation*}
@@
 	\label{pf:first_order_graph_sobolev_seminorm_1}
-	\Ebb\Bigl[f^T \Lap_{n,r} f\Bigr] = \frac{n(n - 1)}{2} \Ebb\biggl[\Bigl(f(X') - f(X)\Bigr)^2 K_r(X',X)\biggr]
+	\Ebb\Bigl[f^T \Lap_{n,r} f\Bigr] = \frac{n(n - 1)}{2} \Ebb\biggl[\Bigl(f(X') - f(X)\Bigr)^2 K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr]
 	\end{equation}
@@
 
-	For the remainder of this proof, we will assume that $f \in C^{\infty}(\Xset)$, which we may do without loss of generality because $C^{\infty}(\Xset)$ is dense in $H^1(\Xset)$.
-Moreover, since $\Xset$ is a Lipschitz domain there exists an extension $g \in C^{\infty}(\Rd)$ of $f$ satisfying
+	For the remainder of this proof, we will assume that $f \in C^{\infty}(\Xset)$, which we may do without loss of generality because $C^{\infty}(\Xset)$ is dense in $H^1(\Xset)$ and the expectation on the right hand side of~\eqref{pf:first_order_graph_sobolev_seminorm_1} is continuous in $\Leb^2(\Xset)$.
+Obviously
 	\begin{equation}
+	\Ebb\biggl[\bigl(f(X') - f(X)\bigr)^2K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr] \leq p_{\max}^2 \int_{\Xset} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K_r(x',x) \,dx' \,dx.
 \label{pf:first_order_graph_sobolev_seminorm_2}
-	|g|_{H^1(\Rd)} \leq C|f|_{H^1(\Xset)} \leq C M,
 	\end{equation}
-	for a constant $C$ which may depend on $\Xset$ and $d$, but does not depend on $M$ or $f$. Since $g(x) = f(x)$ and $p(x) \leq p_{\max}$ for all $x \in \Xset$, 
+	and it remains now to bound the integral. Taking a first-order Taylor expansion of $f$, we get
 	\begin{align}
-	\Ebb\Bigl[\bigl(f(X') - f(X)\bigr)^2K_r(X',X)\Bigr] & = \Ebb\Bigl[\bigl(g(X') - g(X)\bigr)^2K_r(X',X)\Bigr] \nonumber \\
-	& \leq p_{\max}^2 \int_{\Reals^d} \int_{\Reals^d} \bigl(g(x') - g(x)\bigr)^2 K_r(x',x) \,dx' \,dx. \label{pf:first_order_graph_sobolev_seminorm_3}
+	\int_{\Xset} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K_r(x',x) \,dx' \,dx & = \int_{\Xset} \int_{\Xset} \biggl[\int_{0}^{1} \nabla f\bigl(x + t(x' - x)\bigr)^T (x' - x)\,dt\biggr]^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx' \,dx \nonumber \\
+	& \overset{(i)}{\leq} \int_{\Xset} \int_{\Xset} \int_{0}^{1} \biggl[\nabla f\bigl(x + t(x' - x)\bigr)^T (x' - x)\biggr]^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dt\,dx' \,dx \nonumber \\
+	& \overset{(ii)}{=} r^{d + 2} \int_{\Xset} \int_{\mc{Z}_r(x)} \int_{0}^{1} \biggl[\nabla f\bigl(x + trz\bigr)^T z\biggr]^2 K\bigl(\|z\|\bigr) \,dt\,dz \,dx \\
+	&  \overset{(iii)}{\leq} r^{d + 2} \int_{\Xset} \int_{\Reals^d} \int_{0}^{1} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) \,dt \,dz \,d\wt{x} \nonumber \\
+	& = r^{d + 2} \int_{\Xset} \int_{\Reals^d} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) \,dz \,d\wt{x} \label{pf:first_order_graph_sobolev_seminorm_3}
 	\end{align}
-	It remains only to bound the integral. Taking a first-order Taylor expansion of $g$, we get
+	where $(i)$ follows by Jensen's inequality, $(ii)$ follows by substituting $z = (x' - x)/r$ with $\mc{Z}_r(x) := \{z: zr + x \in \Xset\}$ and $(iii)$ by substituting $\wt{x} = x + trz$ and noting that
+	\begin{equation*}
+	x \in \Xset ~~\textrm{and}~~ z \in \mc{Z}_r(x) \Longrightarrow x + trz \in \Xset ~~\textrm{and}~~ z \in \Reals^d.
+	\end{equation*}
+	Now, writing $\bigl(\nabla f(\wt{x}) ^T z\bigr)^2 = \bigl(\sum_{i = 1}^{d} f^{(e_i)}(x) z_{i}\bigr)^2$, expanding the square and integrating, we have that for any $\wt{x} \in \Xset$,
 	\begin{align*}
-	\int_{\Reals^d} \int_{\Reals^d} \bigl(g(x') - g(x)\bigr)^2 K_r(x',x) \,dx \,dx' & = \int_{\Rd} \int_{\Rd} \biggl\{\sum_{\abs{\alpha} = 1}G_{\alpha}(x,x') (x' - x)^{\alpha}\biggr\}^2 K_r(x',x) \,dx' \,dx \\
-	& \overset{(i)}{\leq}  \int_{\Rd} \int_{\Rd} \biggl\{\sum_{\abs{\alpha} = 1}\Bigl[G_{\alpha}(x,x')\Bigr]^2\biggr\} \norm{x' - x}^2  K_r(x',x) \,dx' \,dx \\
-	& \overset{(ii)}{\leq} \int_{\Rd} \int_{\Rd} \biggl\{\int_{0}^{1} \sum_{\abs{\alpha} = 1} \Bigl[g^{(\alpha)}\bigl(x + t(x' - x)\bigr)\Bigr]^2 \,dt \biggr\} \norm{x' - x}^2 K_r(x',x) \,dx' \,dx \\
-	& \overset{(iii)}{=} r^{2 + d} \int_{\Rd} \int_{B(0,1)} \biggl\{\int_{0}^{1} \sum_{\abs{\alpha} = 1} \Bigl[g^{(\alpha)}\bigl(x + trz\bigr)\Bigr]^2 \,dt\biggr\} \norm{z}^2 K\bigl(\norm{z}\bigr)  \,dz \,dx \\
-	& \overset{(v)}{=} r^{2 + d} \int_{B(0,1)} \int_{0}^{1} \biggl\{\int_{\Rd} \sum_{\abs{\alpha} = 1} \Bigl[g^{(\alpha)}\bigl(x + trz\bigr)\Bigr]^2 \,dx \biggr\} \norm{z}^2 K\bigl(\norm{z}\bigr) \,dt \,dz \\
-	& \leq r^{2 + d} |g|_{H^1(\Rd)}.
+	\int_{\Reals^d} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) \,dz & = \sum_{i,j = 1}^{d} f^{(e_i)}(\wt{x}) f^{(e_j)}(\wt{x}) \int_{\Rd} z_iz_jK(\|z\|) \,dz \\
+	& = \sum_{i = 1}^{d} \bigl(f^{(e_i)}(\wt{x})\bigr)^2 \int_{\Rd} z_i^2 K\bigl(\|z\|\bigr) \,dz \\
+	& = \sigma_K \|\nabla f(\wt{x})\|^2
 	\end{align*}
-	In the above manipulations, $(i)$ follows by the Cauchy-Schwarz inequality, $(ii)$ follows by Jensen's inequality, $(iii)$ follows from the change of variables $z = (x' - x)/r$, and $(v)$ follows by Fubini's Theorem.
-Along with~\eqref{pf:first_order_graph_sobolev_seminorm_1}-\eqref{pf:first_order_graph_sobolev_seminorm_3}, this proves Lemma~\ref{lem:first_order_graph_sobolev_seminorm}.
+	and by plugging back into~\eqref{pf:first_order_graph_sobolev_seminorm_3}, we obtain the claimed result.
 \end{proof}

### 2021-02-11 notes/weekly_notes/9.23.20.NotesForWeek.tex
Cleaned up notes folder.

@@
 	\label{asmp:kernel}
-	$K:[0,\infty) \to [0,\infty)$ is a non-increasing function supported on $[0,1]$, and its restriction to $[0,1]$ is Lipschitz with Lipschitz constant $L_K$. Additionally, it is normalized so that
+	$K:[0,\infty) \to [0,\infty)$ is a non-increasing function supported on $[0,1]$, $K(1) > 0$, and its restriction to $[0,1]$ is Lipschitz with Lipschitz constant $L_K$.
+Additionally, it is normalized so that
 	\begin{equation*}
@@
 	\end{equation*}
-	and $K(1/2) > 0$. Let $\sigma_{K} := \frac{1}{d}\int_{\Reals^d} \|z\|^2 K(\|z\|) \,dy$.
+	Let $\sigma_{K} := \frac{1}{d}\int_{\Reals^d} \|z\|^2 K(\|z\|) \,dy$.
 \end{enumerate}
@@
 	\end{equation*}
-	for some constant $A_1(\Xset,d)$. 
+	for some constant $a_1(\Xset,d)$. 
 	\item 
-	\label{asmp:r_small_2} For all $x,x' \in \Xset$, we have that $z_{xx'} = \frac{1}{2}x + \frac{1}{2}x' \in \Xset$.
-	\item 
-	\label{asmp:r_small_3} For any $x \in \Xset$, letting $\mc{B}_x:= (\Xset - x)/r \cap B(0,1)$, we have that $\int_{\mc{B}_x} \psi(\|z\|) \,dz \geq a_1(\Xset,d) \cdot \int_{B(0,1)} \psi(\|z\|) \,dz$.
+	\label{asmp:r_small_3} For any $x \in \Xset$, letting $\mc{B}_x:= (\Xset - x)/r~ \cap~B(0,1)$, we have that $\int_{\mc{B}_x} \psi(\|z\|) \,dz \geq a_1(\Xset,d) \cdot \int_{B(0,1)} \psi(\|z\|) \,dz$.
 \end{enumerate}
@@
 	\label{lem:first_order_graph_sobolev_seminorm_discretized_lb}
-	Suppose $r$ satisfies~\ref{asmp:r_small_2}. For any $u \in \Leb^2(P_n)$, 
+	Suppose $\frac{\wt{\delta}}{r} \leq \min\bigl\{\frac{1}{4(2^{d + 2} - 1)}, \frac{K(1)}{2L_K}\bigr\}$. Then for any $u \in \Leb^2(P_n)$, 
 	\begin{equation*}
@@
 	and
+	\begin{equation}
+	\label{eqn:calder19_ot}
+	\bigl\|\wt{p}_n - p\bigr\|_{\infty} \leq A_1\bigl(\theta + \wt{\delta}\bigr)
+	\end{equation}
+\end{theorem}
+
+The assumptions~\ref{asmp:calder19_1} and~\ref{asmp:calder19_2}, as well as the $\Leb^{\infty}$ bound~\eqref{eqn:calder19_ot}, depend on constants which are not explicitly given.
+To make this dependence more transparent on the geometric parameters of the problem, we restate the Theorem in the flat Euclidean setting; for completeness, we provide a proof in~\textcolor{red}{???}.
+
+\begin{proposition}[Restatement of Proposition~2.11 of \textcolor{red}{(Calder2019)}]
+	\label{prop:optimal_transport}
+	Suppose $p$ satisfies~\ref{asmp:bounded_lipschitz_density}. Then there exists a probability measure $\wt{P}_n$ with density $\wt{p}_n$ such that for any $\wt{\delta} \geq n^{-1/d}$,
 	\begin{equation*}
-	\bigl\|\wt{p}_n - p\bigr\|_{\infty} \leq A_1\bigl(\theta + \wt{\delta}\bigr)
+	\inf_{T: T_{\sharp} \wt{P}_n = P_n} \biggl\{\sup_{x \in \Xset}~\bigl|T(x) - x\bigr|\biggr\} \leq A_2(d,\Xset) \wt{\delta}
 	\end{equation*}
-\end{theorem}
-\textcolor{red}{(TODO)}: Replace this Theorem with it's ``flat'' predecessor.
+	and
+	\begin{equation*}
+	\|\wt{p}_n - p\|_{\infty} \leq A_2(d,\Xset) L_p \wt{\delta} + \theta p_{\max} \leq A_1(\wt{\delta} + \theta)
+	\end{equation*}
+	with probability at least $1 - A_2(d,\Xset) n \cdot \exp\bigl\{-a_2(d,\Xset) p_{\min} n\theta^2\wt{\delta}^d\bigr\}$. 
+\end{proposition}
 
-By Theorem~\ref{thm:calder19_2}, the two energies can be related,
+By assumption $A_1(\wt{\delta} + \theta) \leq \frac{1}{2}$. Therefore Proposition~\ref{prop:optimal_transport} allows to relate the energies $\wt{E}_r(f)$ and $E_r(f)$:
 \begin{equation}
@@
 \end{equation}
-and so can the norms
+as well as the norms $\|f\|_{\Leb^2(P)}$ and $\|f\|_{\Leb^2(P_n)}$:
 \begin{equation}
@@
 \begin{equation*}
-K\biggl(\frac{\|x' - x\|}{\wt{r}}\biggr) \leq K\biggl(\frac{\|x_i - x_j\|}{r}\biggr) + \frac{2L_K\wt{\delta}}{\wt{r}} \cdot \1\biggl\{\frac{\|x' - x\|}{\wt{r}}\biggr\}
+K\biggl(\frac{\|x' - x\|}{\wt{r}}\biggr) \leq K\biggl(\frac{\|x_i - x_j\|}{r}\biggr) + \frac{2L_K\wt{\delta}}{\wt{r}} \cdot \1\biggl\{\frac{\|x' - x\|}{\wt{r}} \leq 1\biggr\}
 \end{equation*}
@@
 \begin{align}
-\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & \leq \frac{1}{n^2\wt{r}^{d + 2}} \sum_{i = 1}^{n} \sum_{j = 1}^{n} \Bigl(u(X_i) - U(X_j)\Bigr)^2 K\biggl(\frac{\|x_i - x_j\|}{r}\biggr) + \frac{2L_K\wt{\delta}}{\wt{r}}\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \nonumber \\
-& \leq \biggl(1 + 4\bigl(2^{d + 2} - 1\bigr)\frac{\wt{\delta}}{r}\biggr) b_r(u) + \frac{2L_K\wt{\delta}}{\wt{r}}\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \label{pf:lem_first_order_graph_sobolev_seminorm_discretized_1}.
+\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & \leq \frac{1}{n^2\wt{r}^{d + 2}} \sum_{i = 1}^{n} \sum_{j = 1}^{n} \Bigl(u(X_i) - u(X_j)\Bigr)^2 K\biggl(\frac{\|x_i - x_j\|}{r}\biggr) + \frac{2L_K\wt{\delta}}{\wt{r}}\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \nonumber \\
+& \leq \biggl(1 + 4\bigl(2^{d + 2} - 1\bigr)\frac{\wt{\delta}}{r}\biggr) b_r(u) + \frac{2L_K\wt{\delta}}{\wt{r}}\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \nonumber \\
+& \leq \biggl(1 + 4\bigl(2^{d + 2} - 1\bigr)\frac{\wt{\delta}}{r}\biggr) b_r(u) + \frac{2L_K\wt{\delta}}{K(1)\wt{r}} \wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u\bigr) \nonumber 
 \end{align}
-We focus now on the factor of $\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,1_{[0,1]})$.
-For any $x$ and $x' \in \Xset$ such that $\|x - x'\| \leq r$, note that by assumption $z_{xx'}:= \frac{1}{2}x + \frac{1}{2}x' \in \Xset \cap B(x,\wt{r}/2) \cap B(x',\wt{r}/2)$.
-Moreover by the triangle inequality
-\begin{equation*}
-\bigl|f(x) - f(x')\bigr|^2 \leq 2\bigl|f(x) - f(z_{xx'})\bigr|^2 + 2\bigl|f(z_{xx'}) - f(x')\bigr|^2
-\end{equation*}
-so that
+or in other words
 \begin{align*}
-\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) & \leq \int_{\Xset} \int_{\Xset \cap B(x,r)} \Bigl[ 2\bigl|f(x) - f(z_{xx'})\bigr|^2 +  2\bigl|f(z_{xx'}) - f(x')\bigr|^2\Bigr] \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx.
+\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & \leq \biggl(1 - \frac{2L_K\wt{\delta}}{K(1)\wt{r}}\biggr)^{-1}\biggl(1 + 4\bigl(2^{d + 2} - 1\bigr)\frac{\wt{\delta}}{r}\biggr) b_r(u) \\
+& \leq \biggl(1 + \frac{2L_K}{K(1)} + 8\bigl(2^{d + 2} - 1\bigr)\biggr) \frac{\wt{\delta}}{r}b_r(u)
 \end{align*}
-Now, by a change of variables along with the Lipschitz property of $p$, we can deduce that
-\begin{align*}
- \int_{\Xset} \int_{\Xset \cap B(x,r)} \bigl|f(x) - f(z_{xx'})\bigr|^2 \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx & \leq 2^d \bigl(1 + L_p\wt{r}\bigr)\int_{\Xset} \int_{\Xset \cap B(x,r/2)} \bigl|f(x) - f(z)\bigr|^2 \wt{p}_n(z) \wt{p}_n(x) \,dz \,dx \\
- & = 2^d \bigl(1 + L_p\wt{r}\bigr) \wt{E}_{\wt{r}/2}\bigl(f; \1_{[0,1]}\bigr),
-\end{align*}
-and therefore by symmetry
-\begin{equation*}
-\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \leq 2^{d + 2} \bigl(1 + L_p\wt{r}\bigr) \wt{E}_{\wt{r}/2}\bigl(f; \1_{[0,1]}\bigr).
-\end{equation*}
-Finally, noting that $\wt{E}_{\wt{r}/2}\bigl(f; \1_{[0,1]}) \leq \frac{1}{K(1/2)} \wt{E}_{\wt{r}}\bigl(f)$, the proof is completed by plugging our upper bound on $\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr)$ back into~\eqref{pf:lem_first_order_graph_sobolev_seminorm_discretized_1}.
+where the second inequality follows from the algebraic identities $(1 - t)^{-1} \leq (1 + 2t)$ for any $0 < t < 1/2$ and $(1 + s)(1 + t) < 1 + 2s + t$ for any $0 < t < 1$ and $s > 0$.
+
 \paragraph{Proof of Lemma~\ref{lem:tau_bound}.}
@@
 Then integrating over $\Xset$ with respect to $p$ yields~\eqref{eqn:smoothening_error_energy}.
+
+\paragraph{Proof of Proposition~\ref{prop:optimal_transport}.}
+We start by defining the density $\wt{p}_n$, which will be piecewise constant over a particular partition $\mc{Q}$ of $\Xset$; specifically if $x \in Q$ for $Q \in \mc{Q}$, then
+\begin{equation}
+\label{pf:prop_optimal_transport_0}
+\wt{p}_n(x) := \frac{P_n(Q)}{\vol(Q)}
+\end{equation}
+where $\vol(\cdot)$ denotes the full-dimensional Lebesgue measure.
+
+We now construct the partition $\mc{Q}$, in progressive degrees of generality on the domain $\Xset.$
+\begin{itemize}
+	\item In the special case of the unit cube $\Xset = (0,1)^d$, the partition will simply be the collection of cubes
+	\begin{equation*}
+	\set{Q_k: k \in [\wt{\delta}^{-1}]^d} 
+	\end{equation*}
+	where $Q_k = \wt{\delta}\Bigl([k_1 - 1,k_1] \otimes \cdots \otimes [k_d - 1,k_d]\Bigr)$ and we assume without loss of generality that $\wt{\delta}^{-1} \in \mathbb{N}$.
+	\item If $\mc{\Xset}$ is an open, connected set with smooth boundary, then by Proposition 3.2 of \textcolor{red}{GarciaTrillos14}, there exist a finite number $N(\Xset) \in \mathbb{N}$ of disjoint polytopes which cover $\Xset$.
+Moreover, letting $U_j$ denote the intersection of the $j$th of these polytopes with $\wb{\Xset}$, this proposition establishes that for each $j$ there exists a bi-Lipschitz homeomorphism $\Phi_j: U_j \to [0,1]^d$.
+We take the collection of
+	\begin{equation*}
+	\mc{Q} = \biggl\{\Phi_j^{-1}(Q_k): j = 1,\ldots,N(\Xset)~~\textrm{and}~~k \in [\wt{\delta}^{-1}]^d\biggr\}
+	\end{equation*}
+	to be our partition. Denote by $L_{\Phi}$ the maximum of the bi-Lipschitz constants of $\Phi_1,\ldots,\Phi_{N(\Xset)}$.
+	\item Finally, in the general case where $\Xset$ is an open, connected set with Lipschitz boundary, then there exists a bi-Lipschitz homeomorphism $\Psi$ between $\Xset$ and a smooth, open, connected set with Lipschitz boundary.
+Letting $\Phi_j$ and $\wt{Q}_{j,k}$ be as before, we take the collection
+	\begin{equation*}
+	\mc{Q} = \biggl\{\wt{Q}_{j,k} = \Bigl(\Psi^{-1} \circ \Phi_j^{-1}\Bigr)(Q_k): j = 1,\ldots,N(\Xset)~~\textrm{and}~~k \in [\wt{\delta}^{-1}]^d\biggr\}
+	\end{equation*}
+	to be our partition. Denote by $L_{\Psi}$ the bi-Lipschitz constant of $\Psi$.
+\end{itemize}
+Let us record a few facts which hold for all $\wt{Q}_{j,k} \in \mc{Q}$, which follow from the bi-Lipschitz properties of $\Phi_j$ and $\Psi$: first that
+\begin{equation}
+\label{pf:prop_optimal_transport_1}
+\diam(\wt{Q}_{j,k}) \leq L_{\Psi} \L_{\Phi} \wt{\delta}
+\end{equation}
+and second that
+\begin{equation}
+\label{pf:prop_optimal_transport_2}
+\vol(\wt{Q}_{j,k}) \geq \biggl(\frac{1}{L_{\Psi} L_{\Phi}}\biggr)^d \wt{\delta}^d.
+\end{equation}
+We now use these facts to show that $\wt{P}_n$ satisfies the claims of Proposition~\ref{prop:optimal_transport}.
+One on the one hand, by the construction~\eqref{pf:prop_optimal_transport_0} clearly there exists a transport map $\wt{T}: \Xset \to \mathbf{X}$ for which every $x \in \wt{Q}_{j,k}$ is moved to a design point $X \in {\bf X} \cap \wt{Q}_{j,k}$, and so by~\eqref{pf:prop_optimal_transport_1}
+\begin{equation*}
+\sup_{x \in \Xset} \abs{\wt{T}(x) - x} \leq  L_{\Psi} \L_{\Phi} \wt{\delta}.
+\end{equation*}
+On the other hand, applying the triangle inequality we have that for $x \in \wt{Q}_{j,k}$
+\begin{align*}
+|\wt{p}_n(x) - p(x)| \leq \biggl|\frac{P_n(\wt{Q}_{j,k}) - P(\wt{Q}_{j,k})}{\vol(\wt{Q}_{j,k})}\biggr| + \frac{1}{\vol(\wt{Q}_{j,k})} \int_{\wt{Q}_{j,k}} |p(x') - p(x)| \,dx 
+\end{align*}
+and using the Lipschitz property of $p$ we find that 
+\begin{equation}
+\label{pf:prop_optimal_transport_3}
+\|\wt{p}_n - p\|_{\Leb^{\infty}} \leq \max_{j,k} \biggl|\frac{P_n(\wt{Q}_{j,k}) - P(\wt{Q}_{j,k})}{\vol(\wt{Q}_{j,k})}\biggr| + L_p L_{\Phi} L_{\Psi} \wt{\delta}
+\end{equation}
+From Hoeffding's inequality and a union bound, we obtain that 
+\begin{align*}
+\mathbb{P}\biggl( \bigl|P_n(\wt{Q}) - P(\wt{Q})\bigr| & \leq \theta P(\wt{Q}) \quad \forall \wt{Q} \in \mc{Q} \biggr) \geq 1 - 2 \sharp(\mc{Q}) \cdot \exp\biggl\{-\frac{\theta^2 n \min \{P(\wt{Q})\}}{3}\biggr\} \\
+& \geq 1 - \frac{2 N(\Xset)}{\wt{\delta}^d} \cdot \exp\biggl\{-\frac{\theta^2 n p_{\min} \wt{\delta}^d }{3\bigl(L_{\Psi} L_{\Phi}\bigr)^d}\biggr\}.
+\end{align*}
+Noting that by assumption $P(\wt{Q}) \leq p_{\max} \vol(\wt{Q})$ and $\wt{\delta}^{-d} \leq n$, the claim follows upon plugging back into~\eqref{pf:prop_optimal_transport_3}, and setting
+\begin{equation*}
+a_1(d, \Xset) := \frac{1}{3\bigl(L_{\Psi} L_{\Phi}\bigr)^d}~~\textrm{and}~~A_1(d,\Xset) := \max \Bigl\{2N(\Xset),L_p L_{\Psi} L_{\Phi} \Bigr\}
+\end{equation*}
+in the statement of the proposition.
 

### 2021-02-11 code/estimation_mse_pipeline.R
Current version of code folder.

@@ -2,13 +2,12 @@ library(RANN)
 library(dplyr)
 library(Matrix)
 library(reshape2)
-library(gridExtra)
 source("sample.R")
 source("graph.R")
 source("estimators.R")
 
 # Please change this line to whichever config you wish to run.
-source("configs/estimation_mse_config5.R")
+source("configs/estimation_mse_config3.R")
 
 #----------------------------------------------------#
 # This is the pipeline for running regression estimation experiments.

### 2021-02-11 code/estimators.R

@@ -89,7 +89,7 @@ initialize_laplacian_smoothing_knn_thetas <- function(sample_X,n){
   thetas <- data.frame(k = numeric(), rho = numeric())
   
   # Choose a range of connectivity parameters.
-  ks <- seq(2, 17, by = 3)
+  ks <- seq(2, 8, by = 2)
   
   # Choose a range of penalty parameters.
   m <- choose_f0_frequency(d)
@@ -113,11 +113,11 @@ initialize_laplacian_smoothing_knn_thetas <- function(sample_X,n){
 }
 
 initialize_knn_thetas <- function(sample_X,n){
-  m <- choose_f0_frequency(d)
-  a <- choose_f0_amplitude(d,m)
-  k_min <- 2
-  k_max <- min(max(round(5 * (d * (a^2/2^d) * m^2 * pi^2)^{-d/(2+d)} * n^{2/(d + 2)}), 50), n/4)
-  ks <- c(2:12, seq(13,k_max,by = 2))
+  # m <- choose_f0_frequency(d)
+  # a <- choose_f0_amplitude(d,m)
+  # k_min <- 2
+  # k_max <- min(max(round(5 * (d * (a^2/2^d) * m^2 * pi^2)^{-d/(2+d)} * n^{2/(d + 2)}), 50), n/4)
+  ks <- c(2:13, seq(14,30,by = 2))
   
   thetas <- data.frame(k = ks)
 }

### 2021-02-11 code/plot.R

@@ -2,7 +2,7 @@ library(ggplot2)
 library(gridExtra)
 
 ### Plot. ###
-save_directory <- "data/20201012191644" # Please change this to whichever directory you prefer.
+save_directory <- "data/AISTATS/cosine/d2_largesample" # Please change this to whichever directory you prefer.
 plot_directory <- file.path(save_directory,"plots")
 dir.create(plot_directory)
 
@@ -88,13 +88,14 @@ fitted_slopes <- numeric(length(methods))
 names(fitted_slopes) <- names(methods)
 for(jj in 1:length(methods))
 {
+  if(names(methods)[jj] == "knn") next
   best_mse <- numeric()
   minimax_mse <- numeric()
   for(ii in 1:length(ns))
   {
     mse_ii_jj <- mse[[ii]][[jj]]
     best_mse[ii] <- min(rowMeans(mse_ii_jj))
-    minimax_mse[ii] <- ns[ii]^{-2/(2 + d)}
+    minimax_mse[ii] <- 4*ns[ii]^{-2/(2 + d)}
   }
   plot_dfs_best_mse[[jj]] <- data.frame(x = ns, y = best_mse, z = minimax_mse) 
   
@@ -103,21 +104,39 @@ for(jj in 1:length(methods))
   log_ns <- log(ns)
   fitted_slopes[jj] <- round( lm(log_best_mse ~ log_ns)$coefficients[2], 2)
   
+  # hack to change names for plotting
+  if(names(plot_dfs_best_mse)[[jj]] == "laplacian_smoothing") names(plot_dfs_best_mse)[[jj]] <- "LS"
   names(plot_dfs_best_mse)[[jj]] <- paste0(names(plot_dfs_best_mse)[[jj]],
-                                           "[Slope =", fitted_slopes[jj],"].")
+                                           " [Slope = ", fitted_slopes[jj],"].")
 }
 
 title <- paste0("d = ", d,". Minimax slope = ", -2,"/", d+2, ".")
 plot_df_best_mse <- bind_rows(plot_dfs_best_mse, .id = "method")
-  TEMP <- 
-    ggplot(data = plot_df_best_mse, aes(x = x)) +
-    coord_trans(x = "log10", y = "log10") + 
-    geom_line(aes(y = y, color = method)) + 
-    geom_point(aes(y = y, color = method)) + 
-    geom_line(aes(y = z)) + 
[…]

### 2021-02-11 code/sample.R

@@ -6,7 +6,6 @@ make_sample_uniform <- function(d)
   }
 }
 
-
 make_cosine_f0 <- function(d,n)
 {
   # Choose m such that the first-order regression error dominates.
@@ -21,6 +20,33 @@ make_cosine_f0 <- function(d,n)
   return(f0)
 }
 
+make_linear_f0 <- function(d,n)
+{
+  #f_0 = b*(X_1 + ... + X_d)
+  
+  # Choose slope
+  b <- choose_linear_f0_slope(d)
+  
+  f0 <- function(x){
+    return(b * sum(x))
+  }
+  return(f0)
+}
+
+choose_linear_f0_slope <- function(d)
+{
+  m <- choose_f0_frequency(d)
+  a <- choose_f0_amplitude(d,m)
+  b <- a*m*pi
+  if(d == 3)
+  {
+    b <- b / 3
+  } else if(d == 4)
+  {
+    b <- b / 4
+  }
+  return(b)
+}
 choose_f0_frequency <- function(d)
 {
   if(d == 1)
@@ -42,8 +68,10 @@ choose_f0_amplitude <- function(d,m)
   if (d <= 3) 
   {
     a <- 3
-  } else {
+  } else if (d == 4){
     a <- 4
+  } else if (d == 5){
+    a <- 6
   }
     
   return(a)

### 2021-02-14 manuscripts/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
Renamed bibliography.
AJG: took another pass over proof, some cosmetic changes to main text.

--- 2020-10-16 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing.tex
@@
 \documentclass[twoside]{article}
-\usepackage{aistats2021}
+\usepackage[accepted]{aistats2021}
 \usepackage{amsmath, amsfonts, amsthm, amssymb}
 \usepackage{graphicx}
-\usepackage[colorlinks]{hyperref}
+\usepackage{hyperref}
+\hypersetup{
+	colorlinks=true,
+	linkcolor=blue,
+	citecolor=blue
+}
 \usepackage[parfill]{parskip}
@@
 
+\runningtitle{Minimax Optimal Laplacian Smoothing}
+
 \twocolumn[
@@
 There are known extremely fast and reliable solvers for exactly this problem (see e.g.
-the seminal papers of \cite{spielman2011,spielman2013,spielman2014}, or the monograph by~\cite{vishnoi2012} and references therein).
+the seminal papers of \citet{spielman2011,spielman2013,spielman2014}, or the monograph by~\citet{vishnoi2012} and references therein).
 	\item \emph{Generality.} Laplacian smoothing is well defined whenever one can associate a graph with some observed responses.
 This generality lends itself to many different data modalities: for instance text and image classification, to name only two \citep{kondor2002, belkin03a,belkin2006}.
-	\item \emph{Weak supervision.} Although we study Laplacian smoothing in the supervised learning setup~\eqref{eqn:signal_plus_noise_model}, it can be easily adapted to the semi-supervised or unsupervised settings \citep{nadler09,slepcev17,dunlop2020,calder2019b}.
+	\item \emph{Weak supervision.} Although we study Laplacian smoothing in the supervised learning setup~\eqref{eqn:signal_plus_noise_model}, it can be easily adapted to the semi-supervised or unsupervised settings \citep{nadler09,slepcev17,calder2019b,dunlop2020}.
 \end{itemize}
@@
 To conduct graph-based inference, the user first builds a neighborhood graph over the random design points---so that $W_{ij}$ is large iff $X_i$ and $X_j$ are close in say Euclidean distance---and then solves e.g.~\eqref{eqn:laplacian_smoothing} or~\eqref{eqn:laplacian_smoothing_test}.
-In this context, various graph-based procedures have been shown to be \emph{consistent}: as $n \to \infty$, they converge to a \emph{continuum} limit (see \citep{belkin07,trillos2018,vonluxburg2008} among others).
+In this context, various graph-based procedures have been shown to be \emph{consistent}: as $n \to \infty$, they converge to a \emph{continuum} limit (see \citet{belkin07,trillos2018,vonluxburg2008} among others).
 However, until recently such statements were not accompanied by error rates, and such error rates as have been proved \citep{lee2016,trillos2020} are not optimal over continuum function spaces, such as H\"{o}lder or Sobolev classes.
@@
 % For an integer $p > 0$, we use $\Leb^p(\Xset)$ to refer to the set of functions $f$ for which the $p$th moment $\norm{f}_{\Leb^p(\Xset)}^p := \int_{\Xset} |f(x)|^p \,dx < \infty$, and $C^p(\Xset)$ to refer to those function which are $p$-times continuously differentiable with bounded $p$th derivative.
-Finally, we use $C,C_1,\ldots$ and $c,c_1,\ldots$ as constants which are independent of $n$ and $f_0$, and for convenience, we will also use asymptotic notation: $a_n \lesssim b_n$ means that $a_n \leq Cb_n$ for some $C > 0$, and $a_n \asymp b_n$ means that $a_n \lesssim b_n$ and $b_n \lesssim a_n$.
+Finally, we use $C,C_1,\ldots$ and $c,c_2,\ldots$ as constants which are independent of $n$ and $f_0$, and for convenience, we will also use asymptotic notation: $a_n \lesssim b_n$ means that $a_n \leq Cb_n$ for some $C > 0$, and $a_n \asymp b_n$ means that $a_n \lesssim b_n$ and $b_n \lesssim a_n$.
 
@@
 \end{equation*}
-For further details regarding Sobolev spaces see e.g.~\citep{evans10,leoni2017}.
+For further details regarding Sobolev spaces see e.g.~\citet{evans10,leoni2017}.
 
@@
 
-Under~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density}, the minimax estimation rate over Sobolev balls is (see e.g. \citep{tsybakov2008_book}),
+Under~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density}, the minimax estimation rate over Sobolev balls of radius $M \geq n^{-1/2}$ is (see e.g. \citet{tsybakov2008_book}),
 \begin{equation}
@@
 \end{equation}
+(Here and throughout, we assume $M \geq n^{-1/2}$, since otherwise the minimax rate is simply the parametric rate $n^{-1}$.)
 
@@
 Typically, the null hypothesis $f_0 = f_0^{\star} \in \mc{F}$ reflects the absence of interesting structure, and $\mc{F} \setminus  \{f_0^{\star}\}$ is a set of smooth departures from this null.
-To fix ideas, as in \cite{ingster2009} we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\star} = 0$ and $\mc{F} = H^1(\Xset,M)$ is a first-order Sobolev ball; this is without loss of generality.
+To fix ideas, as in \citet{ingster2009} we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\star} = 0$ and $\mc{F} = H^1(\Xset,M)$ is a first-order Sobolev ball; this is without loss of generality.
 
@@
 where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation under the regression function $f_0$.\footnote{ Clearly, the minimax critical radius $\epsilon$ depends on $\alpha$.
-However we adopt the typical convention of treating $\alpha$ as a small but fixed positive constant; hence it will not affect the testing error rates, and we suppress it notationally.} See \citep{ingster82, ingster87, ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
+However we adopt the typical convention of treating $\alpha$ as a small but fixed positive constant; hence it will not affect the testing error rates, and we suppress it notationally.} See \citet{ingster82, ingster87, ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
 
@@
 Suppose that that the neighborhood graph $G_{n,r}$ is computed with a radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_estimation},  and the Laplacian smoothing estimator $\wh{f}$ with $\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$.
-With probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c M^{d/(2d + 4)} n^{d/(2+d)})$, it holds that
+With probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c \min\{M^{d/(2d + 4)} n^{d/(2+d)},n\})$, it holds that
 	\begin{equation}
@@
 	\item It is possible to extend $\wh{f}$ to be defined on all of $\Xset$, and evaluate the error of such an extension in $\Leb^2(\Xset)$ norm.
-Assuming that the extension of $\wh{f}$ and the regression function $f_0$ are both suitably smooth, tools from empirical process theory (Chapter 14 of \cite{wainwright2019}) or approximation theory (Section 15.5 of \citep{johnstone2011}) should guarantee that the $\Leb^2(\Xset)$ error is not too much greater than its in-sample counterpart.
+Assuming that the extension of $\wh{f}$ and the regression function $f_0$ are both suitably smooth, tools from empirical process theory (Chapter 14 of \citet{wainwright2019}) or approximation theory (Section 15.5 of \citet{johnstone2011}) should guarantee that the $\Leb^2(\Xset)$ error is not too much greater than its in-sample counterpart.
 In fact, as we show in our supplementary material, if $f_0$ is Lipschitz smooth and we extend $\wh{f}$ to be piecewise constant over the Voronoi tesselation induced by $X_1,\ldots,X_n$, the out-of-sample error $\|f_0 - \wh{f}\|_{\Leb^2(X)}$ is within a negligible factor of the in-sample error $\|f_0 - \wh{f}\|_n$, but we leave analysis of the Sobolev case to future work.
@@
 When $d = 4$, by setting $r \asymp (\log n/n)^{1/4}$ our analysis results in an upper bound on the mean squared error within a $(\log n)^{1/3}$ factor of the minimax risk, but when $d \geq 5$ our analysis yields upper bounds that are much worse than the minimax rate.
-This mirrors the conclusions of~\cite{sadhanala16}, who investigate estimation rates of Laplacian smoothing over the $d$-dimensional grid graph.
-\cite{sadhanala16} argue that their analysis is tight, and that it is the estimator, rather than the analysis, which is deficient when $d \geq 5$.
+This mirrors the conclusions of~\citet{sadhanala16}, who investigate estimation rates of Laplacian smoothing over the $d$-dimensional grid graph.
+\citet{sadhanala16} argue that their analysis is tight, and that it is the estimator, rather than the analysis, which is deficient when $d \geq 5$.
 Formally proving such a claim turns out to be harder in the random design setting than in the fixed design setting, and we leave it for future work.
@@
 \end{enumerate}
-Note that when $d < 4$, if $M$ is not too large there always exists a choice of $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}.
+Note that when $d < 4$, if $M$ is not too large there always exists a choice of $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}. Put
+\begin{equation*}
+M_{\max}(d) := 
+\begin{cases*}
+n^{1/8},& \textrm{$d = 1$}\\
+n^{(4 - d)/(4d)},& \textrm{$d \geq 2$}.
+\end{cases*}
+\end{equation*}
 \begin{theorem}
 	\label{thm:laplacian_smoothing_testing}
-	Let $f_0 \in H^1(\Xset,M)$ for $M \leq n^{(4 - d)/(4 + d)}$ and $d < 4$.
+	Let $f_0 \in H^1(\Xset,M)$ for $M \leq M_{\max}(d)$ and $d < 4$.
 Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}, and the Laplacian smoothing test $\wh{\varphi}$ with $\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$, and $b \geq 1$.
@@
 	& \Ebb_{f_0}\Bigl[1 - \wh{\varphi} \Bigr] \leq \\
-	& C\biggl(\frac{1}{b}\Bigl[1 + n^{-d(4 + d)}M^{-2d/(4 + d)}\Bigr] + C_1n\exp\bigl(-c_1nr^d\bigr)\biggr).
+	& C\biggl(\frac{1}{b}\Bigl[1 + n^{-d/(4 + d)}M^{-2d/(4 + d)}\Bigr] + n\exp\bigl(-c_1nr^d\bigr)\biggr).
 	\end{align*}
@@
 Proving tight lower bounds over such classes is non-trivial, and to the best of our knowledge such an analysis remains outstanding.
-On the other hand, if we explicitly assume that $f_0 \in \Leb^4(\Xset,M)$, then \cite{guerre02} show that the testing problem is characterized by the dimension-free lower bound $\epsilon^{2}(\Leb^4(\Xset,M)) \gtrsim n^{-1/2}$.
+On the other hand, if we explicitly assume that $f_0 \in \Leb^4(\Xset,M)$, then \citet{guerre02} show that the testing problem is characterized by the dimension-free lower bound $\epsilon^{2}(\Leb^4(\Xset,M)) \gtrsim n^{-1/2}$.
 Moreover, by training $\wh{f}$ to the interpolation limit, i.e.
@@
 First, we should make clear why this difference in performance is so surprising.
-As mentioned previously, the penalties in~\eqref{eqn:laplacian_smoothing} and~\eqref{eqn:thin_plate_spline} can be closely tied together: \cite{bousquet03} show that for any $f \in C^2(\Xset)$,
+As mentioned previously, the penalties in~\eqref{eqn:laplacian_smoothing} and~\eqref{eqn:thin_plate_spline} can be closely tied together: \citet{bousquet03} show that for any $f \in C^2(\Xset)$,
 \begin{equation}
@@
 \lim \frac{1}{n^2 r^{d + 2}} f^T \Lap_{n,r} f & = \int_{\Xset} f(x) \cdot \Delta_Pf(x) p(x) \,dx \\
-& = \int_{\Xset} \|\nabla f(x)\|^2 p^2(x) \,dx
+& = \int_{\Xset} \|\nabla f(x)\|^2 p^2(x) \,dx.
 \end{aligned}
 \end{equation}
-in the above, the limit is as $n \to \infty$ and $r \to 0$, $\Delta_P$ is the (weighted) Laplace-Beltrami operator
+In the above, the limit is as $n \to \infty$ and $r \to 0$, $\Delta_P$ is the (weighted) Laplace-Beltrami operator
 \begin{equation*}
-\Delta_Pf(x) := -\frac{1}{p} \dive\bigl(p^2\nabla f(x))
+\Delta_Pf := -\frac{1}{p} \dive\bigl(p^2\nabla f)
 \end{equation*}
-and the second equality follows from integration by parts.\footnote{Assuming $f$ satisfies Dirichlet boundary conditions.} To be clear, this argument does not formally imply that $\wh{f}$ and $\wt{f}$ are close---note that~\eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:thin_plate_spline} is $H^1(\Xset)$---but it does seem to suggest that the two estimators should behave somewhat similarly.
+and the second equality follows from integration by parts.\footnote{Assuming $f$ satisfies e.g.
+Dirichlet boundary conditions.} To be clear, this argument does not formally imply that $\wh{f}$ and $\wt{f}$ are close---note that~\eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:thin_plate_spline} is $H^1(\Xset)$---but it does seem to suggest that the two estimators should behave somewhat similarly.
 
@@
 This is meant in various related senses.
-From the Sobolev embedding theorem \cite{evans10}, we know that $H^1(\Xset)$ does not continuously embed into any H\"{o}lder space, and that $H^1(\Xset)$ is not an Reproducing Kernel Hilbert Space.
+From the Sobolev embedding theorem \citep{evans10}, we know that $H^1(\Xset)$ does not continuously embed into any H\"{o}lder space, and that $H^1(\Xset)$ is not a Reproducing Kernel Hilbert Space.
 More relevant to this context, by placing ``bump functions'' of arbitrarily small radius around each of the design points $X_1,\ldots,X_n$, one can interpolate the responses $Y_1,\ldots,Y_n$ in such a manner that forces the criterion in~\eqref{eqn:thin_plate_spline} to 0 \citep{green93}; the resulting estimator will clearly be inconsistent.
@@
 	\label{eqn:graph_sobolev_seminorm}
-	f^T \Lap_{n,r} f \leq \frac{(1 + L_p)^2 \sigma_K}{\delta} n^2 r^{d + 2} |f|_{H^1(X)}^2.
+	f^T \Lap_{n,r} f \leq \frac{C_2}{\delta} n^2 r^{d + 2} |f|_{H^1(X)}^2.
 	\end{equation}
@@
 	\label{eqn:neighborhood_eigenvalue}
-	c_2A_{n,r}(k) \leq \lambda_k \leq C_2A_{n,r}(k)~~\textrm{for all $2 \leq k \leq n$}
+	c_3A_{n,r}(k) \leq \lambda_k \leq C_3A_{n,r}(k)~~\textrm{for all $2 \leq k \leq n$}
 	\end{equation}
@@
 
-On the other hand, a theory has been developed~\citep{belkin03,belkin05,niyogi2013} establishing the the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+On the other hand, a theory has been developed~\citep{belkin03,belkin05,niyogi2013} establishing that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
 In this section, we contribute to this line of work by showing that under the manifold hypothesis, Laplacian smoothing achieves the sharper minimax rates over $H^1(\Xset)$.
@@
 	\begin{equation*}
-	C_3\biggl(\frac{\log n}{n}\biggr)^{1/m} \leq r \leq n^{\frac{-3}{(4 + 2m)}} M^{\frac{(m - 4)}{(4 + 2m)}} \wedge c_3.
+	C_4\biggl(\frac{\log n}{n}\biggr)^{1/m} \leq r \leq n^{\frac{-3}{(4 + 2m)}} M^{\frac{(m - 4)}{(4 + 2m)}} \wedge c_4.
 	\end{equation*}
@@
 Suppose that that the neighborhood graph $G_{n,r}$ is computed with a radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_estimation_manifold},  and the Laplacian smoothing estimator $\wh{f}$ with $\rho = M^{-4/(2 + m)} (nr^{m + 2})^{-1} n^{-2/(2 + m)}$.
-With probability at least $1 - \delta -  C_4n\exp(-c_4nr^m) - \exp(-c M^{m/(2m + 4)} n^{m/(2+m)})$, it holds that
+With probability at least $1 - \delta -  Cn\exp(-cnr^m) - \exp(-c M^{m/(2m + 4)} n^{m/(2+m)})$, it holds that
 	\begin{equation*}
@@
 	\begin{equation*}
-	C_3\biggl(\frac{\log n}{n}\biggr)^{1/m} \leq r \leq M^{\frac{(m - 8)}{8 + 2m}}n^{\frac{m - 20}{32 + 8m}} \wedge c_3.
+	C_4\biggl(\frac{\log n}{n}\biggr)^{1/m} \leq r \leq M^{\frac{(m - 8)}{8 + 2m}}n^{\frac{m - 20}{32 + 8m}} \wedge c_4.
 	\end{equation*}
@@
 	\label{thm:laplacian_smoothing_testing_manifold}
-	Let $f_0 \in H^1(\Xset,M)$ for $M \leq n^{(4 - m)/(4 + m)}$ and $m < 4$.
+	Let $f_0 \in H^1(\Xset,M)$ for $M \leq M_{\max}(m)$ and $m < 4$.
 Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing_manifold}, and the Laplacian smoothing test $\wh{\varphi}$ with $\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$, and $b \geq 1$.
@@
 	& \Ebb_{f_0}\Bigl[1 - \wh{\varphi} \Bigr] \leq \\
-	& C\biggl(\frac{1}{b}\Bigl[1 + n^{-m(4 + m)}M^{-2m/(4 + m)}\Bigr] + C_4n\exp\bigl(-c_4nr^m\bigr)\biggr).
+	& C\biggl(\frac{1}{b}\Bigl[1 + n^{-m/(4 + m)}M^{-2m/(4 + m)}\Bigr] + n\exp\bigl(-cnr^m\bigr)\biggr).
 	\end{align*}
@@
 There are still several extensions we feel are worth pursuing.
-In practice, it is more common to use $k$NN graphs than neighborhood graphs, due to the guaranteed connectivity and sparsity of the former; we suspect that by building on the work of~\cite{calder2019}, one can show that our main results all hold when the $k$NN graph is used.
+In practice, it is more common to use $k$NN graphs than neighborhood graphs, due to the guaranteed connectivity and sparsity of the former; we suspect that by building on the work of~\citet{calder2019}, one can show that our main results all hold when the $k$NN graph is used.
 In another direction, one can also generalize Laplacian smoothing by replacing the penalty $f^T \Lap_{n,r} f$ with $f^T \Lap_{n,r}^k f$ for some $k > 1$.
@@
 \bibliographystyle{plainnat}
-\bibliography{../../../graph_testing_bibliography}
+\bibliography{../../../graph_regression_bibliography}
 

### 2021-02-14 manuscripts/laplacian_smoothing/AISTATS/supplement_minimax_optimal_laplacian_smoothing.tex
Renamed bibliography.
AJG: took another pass over proof, some cosmetic changes to main text.

--- 2020-10-22 notes/manuscripts/AISTATS/supplement_minimax_optimal_laplacian_smoothing.tex
@@
 \documentclass[twoside]{article}
-\usepackage{aistats2021}
+\usepackage[accepted]{aistats2021}
 \usepackage{amsmath, amsfonts, amsthm, amssymb}
@@
 \usepackage{xr-hyper}
-\usepackage[colorlinks]{hyperref}
+\usepackage{hyperref}
+\hypersetup{
+	colorlinks=true,
+	linkcolor=blue,
+	citecolor=blue
+}
 \externaldocument{minimax_optimal_laplacian_smoothing}
@@
 \label{eqn:ls_G}
-\wh{f} := \argmin_{f \in \Reals^n} \biggl\{ \sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap f \biggr\} = (\rho \Lap + \Id)^{-1}{\bf Y}.
+\wh{f} := \argmin_{f \in \Reals^n} \biggl\{ \sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^{\top} \Lap f \biggr\} = (\rho \Lap + \Id)^{-1}{\bf Y}.
 \end{equation*}
@@
 \end{equation*}
-We note that in this section, many of the derivations involved in upper bounding the estimation error of $\wh{f}$ are similar to those of \citep{sadhanala16}, with the difference being that we seek bounds in high probability rather than in expectation.
+We note that in this section, many of the derivations involved in upper bounding the estimation error of $\wh{f}$ are similar to those of \cite{sadhanala16}, with the difference being that we seek bounds in high probability rather than in expectation.
 We keep the work here self-contained for purposes of completeness.
@@
 In  Lemma~\ref{lem:linear_smoother_fixed_graph_estimation} we upper bound the error $\frac{1}{n}\|\wc{f} - f_0\|_2^2$ as a function of the eigenvalues of $S$.
-Let $\lambdavec(S) = (\lambda_1(S),\ldots,\lambda_n(S)) \in \Reals^n$ denote these eigenvalues, and let $v_k(S)$ denote the corresponding unit-norm eigenvectors, so that $S = \sum_{k = 1}^{n} \lambda_k(S) \cdot v_k(S) v_k(S)^T$.
-Denote $Z_k = v_k(S)^T \varepsilon$, and observe that ${\bf Z} = (Z_1,\ldots,Z_n) \sim N(0,\Id)$.
+Let $\lambdavec(S) = (\lambda_1(S),\ldots,\lambda_n(S)) \in \Reals^n$ denote these eigenvalues, and let $v_k(S)$ denote the corresponding unit-norm eigenvectors, so that $S = \sum_{k = 1}^{n} \lambda_k(S) \cdot v_k(S) v_k(S)^{\top}$.
+Denote $Z_k = v_k(S)^{\top} \varepsilon$, and observe that ${\bf Z} = (Z_1,\ldots,Z_n) \sim N(0,\Id)$.
 
@@
 
-In Lemma~\ref{lem:linear_smoother_fixed_graph_testing}, we upper bound the error of a test involving the statistic $\|\wc{f}\|_2^2 = {\bf Y}^T S^2 {\bf Y}$.
+In Lemma~\ref{lem:linear_smoother_fixed_graph_testing}, we upper bound the error of a test involving the statistic $\|\wc{f}\|_2^2 = {\bf Y}^{\top} S^2 {\bf Y}$.
 \begin{lemma}
 	\label{lem:linear_smoother_fixed_graph_testing}
-	Let $\wc{T} = Y^T S^2 Y$ for a square, symmetric matrix $S \in \Reals^{n \times n}$. Define the threshold $\wc{t}_b$ to be 
+	Let $\wc{T} = Y^{\top} S^2 Y$ for a square, symmetric matrix $S \in \Reals^{n \times n}$. Define the threshold $\wc{t}_b$ to be 
 	\begin{equation}
@@
 		\label{eqn:linear_smoother_fixed_graph_testing_critical_radius}
-		f_0^T S^2 f_0 \geq 4b \norm{\lambdavec(S)}_4^2,
+		f_0^{\top} S^2 f_0 \geq 4b \norm{\lambdavec(S)}_4^2,
 		\end{equation}
@@
 \end{align*}
-Writing $\norm{S\varepsilon}_2^2 = \sum_{k = 1}^{n} \lambda_k(S)^2 Z_k^2$, the claim follows from the result of \cite{laurent00} on concentration of $\chi^2$-random variables, which for completeness we restate in Lemma~\ref{lem:chi_square_bound}.
+Writing $\norm{S\varepsilon}_2^2 = \sum_{k = 1}^{n} \lambda_k(S)^2 Z_k^2$, the claim follows from the result of \citet{laurent00} on concentration of $\chi^2$-random variables, which for completeness we restate in Lemma~\ref{lem:chi_square_bound}.
 To be explicit, taking $t = \norm{\lambdavec(S)}_2^2$ in Lemma~\ref{lem:chi_square_bound} completes the proof of Lemma~\ref{lem:linear_smoother_fixed_graph_estimation}.
@@
 
-\textit{Mean.} We make use of the eigendecomposition $S = \sum_{k = 1}^{n} \lambda_k(S) \cdot v_k(S) v_k(S)^T$ to obtain
+\textit{Mean.} We make use of the eigendecomposition $S = \sum_{k = 1}^{n} \lambda_k(S) \cdot v_k(S) v_k(S)^{\top}$ to obtain
 \begin{equation}
@@
 \begin{aligned}
-\wc{T} & = f_0^T S^2 f_0 + 2 f_0^T S^2 \varepsilon + \varepsilon^T S^2 \varepsilon \\
-& = f_0^T S^2 f_0 + 2 f_0^T S^2 \varepsilon + \sum_{k = 1}^{n}  \bigl(\lambda_k(S)\bigr)^2 (\varepsilon^T v_k(S))^2 \\
-& = f_0^T S^2 f_0 + 2 f_0^T S^2 \varepsilon + \sum_{k = 1}^{n}  \bigl(\lambda_k(S)\bigr)^2 Z_k^2,
+\wc{T} & = f_0^{\top} S^2 f_0 + 2 f_0^{\top} S^2 \varepsilon + \varepsilon^{\top} S^2 \varepsilon \\
+& = f_0^{\top} S^2 f_0 + 2 f_0^{\top} S^2 \varepsilon + \sum_{k = 1}^{n}  \bigl(\lambda_k(S)\bigr)^2 (\varepsilon^{\top} v_k(S))^2 \\
+& = f_0^{\top} S^2 f_0 + 2 f_0^{\top} S^2 \varepsilon + \sum_{k = 1}^{n}  \bigl(\lambda_k(S)\bigr)^2 Z_k^2,
 \end{aligned}
@@
 \label{pf:linear_smoother_fixed_graph_testing_mean}
-\Ebb_{f_0}\bigl[\wc{T}\bigr] = f_0^T S^2 f_0 + \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^2.
+\Ebb_{f_0}\bigl[\wc{T}\bigr] = f_0^{\top} S^2 f_0 + \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^2.
 \end{equation}
@@
 \label{pf:linear_smoother_fixed_graph_testing_var}
-\Var_{f_0}\bigl[\wc{T}\bigr] \leq 8 f_0^T S^4 f_0 + 4 \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^4.
+\Var_{f_0}\bigl[\wc{T}\bigr] \leq 8 f_0^{\top} S^4 f_0 + 4 \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^4.
 \end{equation}
@@
 & \overset{(i)}{\leq} \Pbb_{f_0}\Bigl(\Bigl|\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr| \geq \Bigl|\wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr|\Bigr) \leq \frac{Var_{f_0}\bigl[\wc{T}\bigr]}{\bigl(\wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\bigr)^2} \\
-& \overset{(ii)}{\leq} 4 \frac{\Var_{f_0}\bigl[\wc{T}\bigr]}{(f_0^T S^2 f_0)^2} \\
-& \overset{(iii)}{\leq} \frac{32}{f_0^T S^2 f_0} + \frac{4}{b^2} \\
+& \overset{(ii)}{\leq} 4 \frac{\Var_{f_0}\bigl[\wc{T}\bigr]}{(f_0^{\top} S^2 f_0)^2} \\
+& \overset{(iii)}{\leq} \frac{32}{f_0^{\top} S^2 f_0} + \frac{4}{b^2} \\
 & \overset{(iv)}{\leq} \frac{8}{b} \Biggl(\sum_{k = 1}^{n}\bigl(\lambda_k(S)\bigr)^4\Biggr)^{-1/2} + \frac{4}{b^2}
@@
 \begin{equation*}
-\wc{t}_b - \Ebb_{f_0}[\wc{T}] = f_0^T S^2 f_0 - 2b\|{\bf \lambda}(S)\|_4^2 \geq \frac{1}{2}f_0^T S^2 f_0
+\wc{t}_b - \Ebb_{f_0}[\wc{T}] = f_0^{\top} S^2 f_0 - 2b\|{\bf \lambda}(S)\|_4^2 \geq \frac{1}{2}f_0^{\top} S^2 f_0
 \end{equation*}
@@
 	\label{eqn:ls_fixed_graph_estimation_prob}
-	\frac{1}{n}\bigl\|\wh{f} - f_0\bigr\|_2^2 \leq \frac{2\rho}{n} \bigl(f_0^T \Lap f_0\bigr) + \frac{10}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k} + 1\bigr)^2}
+	\frac{1}{n}\bigl\|\wh{f} - f_0\bigr\|_2^2 \leq \frac{2\rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{10}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k} + 1\bigr)^2}
 	\end{equation}
@@
 		\label{eqn:ls_fixed_graph_testing_critical_radius}
-		\frac{1}{n}\norm{f_0}_2^2 \geq \frac{2 \rho}{n} \bigl(f_0^T \Lap f_0\bigr) + \frac{4b}{n} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{1/2}
+		\frac{1}{n}\norm{f_0}_2^2 \geq \frac{2 \rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{4b}{n} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{1/2}
 		\end{equation}
@@
 \paragraph{Proof of Lemma~\ref{lem:ls_fixed_graph_estimation}.}
-Letting $\wh{S} = (\Id + \rho \Lap)^{-1}$, the estimator $\wh{f} = \wh{S}Y$, and
+Let $\wh{S} = (\Id + \rho \Lap)^{-1}$, the estimator $\wh{f} = \wh{S}Y$, and
 \begin{equation*}
@@
 \begin{aligned}
-\bigl\|(\wh{S} - I)f_0\bigr\|_2^2 & = f_0^T \Lap^{1/2} \Lap^{-1/2}\bigl(\wh{S} - \Id\bigr) \Lap^{-1/2} \Lap^{1/2} f_0 \\
-& \leq f_0^T \Lap f_0 \cdot \lambda_n\Bigl(\Lap^{-1/2}\bigl(\wh{S} - \Id\bigr)\Lap^{-1/2}\Bigr) \\
-& = f_0^T \Lap f_0 \cdot \max_{k \in [n]} \biggl\{\frac{1}{\lambda_k} \Bigl(1 - \frac{1}{\rho\lambda_k + 1}\Bigr) \biggr\} \\
-& = f_0^T \Lap f_0 \cdot \rho.
+\bigl\|(\wh{S} - I)f_0\bigr\|_2^2 & = f_0^{\top} \Lap^{1/2} \Lap^{-1/2}\bigl(\wh{S} - \Id\bigr) \Lap^{-1/2} \Lap^{1/2} f_0 \\
+& \leq f_0^{\top} \Lap f_0 \cdot \lambda_n\Bigl(\Lap^{-1/2}\bigl(\wh{S} - \Id\bigr)\Lap^{-1/2}\Bigr) \\
+& = f_0^{\top} \Lap f_0 \cdot \max_{k \in [n]} \biggl\{\frac{1}{\lambda_k} \Bigl(1 - \frac{1}{\rho\lambda_k + 1}\Bigr) \biggr\} \\
+& = f_0^{\top} \Lap f_0 \cdot \rho.
 \end{aligned}
@@
 In the above, we have written $\Lap^{-1/2}$ for the square root of the pseudoinverse of $\Lap$, and the maximum is over all indices $k$ such that $\lambda_k > 0$.
+The claim of the Lemma then follows from Lemma~\ref{lem:linear_smoother_fixed_graph_estimation}.
 
 \paragraph{Proof of Lemma~\ref{lem:ls_fixed_graph_testing}.}
-Let $\wh{S} := (\rho \Lap + \Id)^{-1}$, and note that $\wh{T} = \frac{1}{n}{\bf Y}^T \wh{S}^2 {\bf Y}$.
+Let $\wh{S} := (\rho \Lap + \Id)^{-1}$, and note that $\wh{T} = \frac{1}{n}{\bf Y}^{\top} \wh{S}^2 {\bf Y}$.
 The bound on Type I error~\eqref{eqn:ls_fixed_graph_testing_typeI} follows immediately from \eqref{eqn:linear_smoother_fixed_graph_testing_typeI}.
-To establish the bound on Type II error, we must lower bound $f_0^T \wh{S}^2 f_0$. We first note that by assumption~\eqref{eqn:ls_fixed_graph_testing_critical_radius},
-\begin{align*}
-f_0^{T} \wh{S}^2 f_0 & = \bigl\|f_0\bigr\|_2^2 - f_0^T(I - \wh{S}^2)f_0 \\
-& \geq 2 \rho \bigl(f_0^T \Lap f_0\bigr) - f_0^T\bigl(I - \wh{S}^2\bigr)f_0 + 4b \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2}.
-\end{align*}
-Upper bounding $f_0^T\bigl(I - \wh{S}^2\bigr)f_0$ as follows:
+To establish the bound on Type II error, we must lower bound $f_0^{\top} \wh{S}^2 f_0$. We first note that by assumption~\eqref{eqn:ls_fixed_graph_testing_critical_radius},
+\begin{align*}
+f_0^{\top} \wh{S}^2 f_0 & = \bigl\|f_0\bigr\|_2^2 - f_0^{\top}(I - \wh{S}^2)f_0 \\
+& \geq 2 \rho \bigl(f_0^{\top} \Lap f_0\bigr) - f_0^{\top}\bigl(I - \wh{S}^2\bigr)f_0 + 4b \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2}.
+\end{align*}
+Upper bounding $f_0^{\top}\bigl(I - \wh{S}^2\bigr)f_0$ as follows:
 \begin{equation*}
 \begin{aligned}
-f_0^T \Bigl(\Id - \wh{S}^2\Bigr) f_0  & = f_0^T \Lap^{1/2} \Lap^{-1/2}\Bigl(\Id - \wh{S}^2\Bigr) \Lap^{-1/2} \Lap^{1/2} f_0 \\ 
-& \leq f_0^T \Lap f_0 \cdot  \lambda_{n}\biggl(\Lap^{-1/2}\Bigl(\Id - \wh{S}^2\Bigr) \Lap^{-1/2}\biggr) \\ 
-& \overset{(i)}{=}  f_0^T \Lap f_0 \cdot \max_{k} \biggl\{ \frac{1}{\lambda_k} \Bigl(1 - \frac{1}{(\rho \lambda_k + 1)^2}\Bigr) \biggr\} \\
-& \overset{(ii)}{\leq} f_0^T \Lap f_0 \cdot 2\rho,
+f_0^{\top} \Bigl(\Id - \wh{S}^2\Bigr) f_0  & = f_0^{\top} \Lap^{1/2} \Lap^{-1/2}\Bigl(\Id - \wh{S}^2\Bigr) \Lap^{-1/2} \Lap^{1/2} f_0 \\ 
+& \leq f_0^{\top} \Lap f_0 \cdot  \lambda_{n}\biggl(\Lap^{-1/2}\Bigl(\Id - \wh{S}^2\Bigr) \Lap^{-1/2}\biggr) \\ 
+& \overset{(i)}{=}  f_0^{\top} \Lap f_0 \cdot \max_{k} \biggl\{ \frac{1}{\lambda_k} \Bigl(1 - \frac{1}{(\rho \lambda_k + 1)^2}\Bigr) \biggr\} \\
+& \overset{(ii)}{\leq} f_0^{\top} \Lap f_0 \cdot 2\rho,
 \end{aligned}
 \end{equation*}
----where in the above the maximum is over all indices $k$ such that $\lambda_k > 0$, and the last inequality follows from the basic algebraic identity $1 - 1/(1 + s x)^2 \leq 2 s x$ for any $x, s > 0$---we deduce that
-\begin{equation*}
-f_0^{T} \wh{S}^2 f_0 \geq 2b \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{-1/2}.
+---where in the above the maximum is over all indices $k$ such that $\lambda_k > 0$, and the last inequality follows from the basic algebraic identity $1 - 1/(1 + x)^2 \leq 2 x$ for any $x > 0$---we deduce that
+\begin{equation*}
+f_0^{\top} \wh{S}^2 f_0 \geq 2b \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{-1/2}.
 \end{equation*} 
@@
 \label{sec:graph_sobolev_seminorm}
-In this section, we prove Lemma~\ref{lem:graph_sobolev_seminorm}, which states an upper bound on $f^T \Lap_{n,r} f$ holds when $f$ is bounded in Sobolev norm.
-We also establish stronger bounds in the case when $f$ has a bounded Lipschitz constant; this latter result justifies the remark one of our remarks after Theorem~\ref{thm:laplacian_smoothing_estimation1}.
+In this section, we prove Lemma~\ref{lem:graph_sobolev_seminorm}, which states an upper bound on $f^{\top} \Lap_{n,r} f$ that holds when $f$ is bounded in Sobolev norm.
+We also establish stronger bounds in the case when $f$ has a bounded Lipschitz constant; this latter result justifies one of our remarks after Theorem~\ref{thm:laplacian_smoothing_estimation1}.
 
@@
 \begin{equation*}
-\Ebb\Bigl[f^T \Lap_{n,r} f\Bigr] \leq (1 + L_pr)^2 \sigma_K n^2 r^{d + 2} |f|_{H^1(\Xset)}^2
-\end{equation*}
-whence the claim follows immediately by Markov's inequality (recall that $\Lap_{n,r}$ is positive semi-definite, and therefore $f^T \Lap_{n,r} f$ is a non-negative random variable).
+\Ebb\Bigl[f^{\top} \Lap_{n,r} f\Bigr] \leq C_2 n^2 r^{d + 2} \|f\|_{H^1(\Xset)}^2,
+\end{equation*}
+whence the claim follows immediately by Markov's inequality (recall that $\Lap_{n,r}$ is positive semi-definite, and therefore $f^{\top} \Lap_{n,r} f$ is a non-negative random variable).
 
@@
 \begin{equation*}
-f^T \Lap_{n,r} f = \frac{1}{2}\sum_{i, j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 W_{ij},
+f^{\top} \Lap_{n,r} f = \frac{1}{2}\sum_{i, j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 W_{ij},
 \end{equation*}
@@
 \label{pf:first_order_graph_sobolev_seminorm_1}
-\Ebb\Bigl[f^T \Lap_{n,r} f\Bigr] = \frac{n(n - 1)}{2} \Ebb\biggl[\Bigl(f(X') - f(X)\Bigr)^2 K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr]
+\Ebb\Bigl[f^{\top} \Lap_{n,r} f\Bigr] = \frac{n(n - 1)}{2} \Ebb\biggl[\Bigl(f(X') - f(X)\Bigr)^2 K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr]
 \end{equation}
@@
 
-For the remainder of this proof, we will assume that $f \in C^{\infty}(\Xset)$.
-We may do so without loss of generality: $C^{\infty}(\Xset)$ is dense in $H^1(\Xset)$ and the expectation on the right hand side of~\eqref{pf:first_order_graph_sobolev_seminorm_1} is continuous in $H^1(\Xset)$.
+Take $\Omega$ to be an arbitrary bounded open set such that $B(x,r) \subset \Omega$ for all $x \in \mc{X}$.
+For the remainder of this proof, we will assume that (i) $f \in H^1(\Omega)$ and additionally (ii) $\|f\|_{H^1(\Omega)} \leq C_5 \|f\|_{H^1(\mc{X})}$ for a constant $C_5$ that does not depend on $f$.
+This is without loss of generality, since by Theorem~1 of~\citet{evans10} there exists an extension operator $E: H^1(\mc{X}) \to H^1(\Omega)$ for which the extension $Ef$ satisfies both (i) and (ii).
+Additionally, we will assume $f \in C^{\infty}(\Omega)$.
+Again, this is without loss of generality, as $C^{\infty}(\Omega)$ is dense in $H^1(\Omega)$ and the expectation on the right hand side of~\eqref{pf:first_order_graph_sobolev_seminorm_1} is continuous in $H^1(\Omega)$.
+The reason for dealing with a smooth extension $f \in C^{\infty}(\Omega)$ is so that we can make sense of the following equality for any $x$ and $x'$ in $\mc{X}$:
+\begin{equation}
+\label{pf:first_order_graph_sobolev_seminorm_1.5}
+f(x') - f(x) = \int_{0}^{1} \nabla f\bigl(x + t(x' - x)\bigr)^{\top} (x' - x) \,dt.
+\end{equation}
+
 Obviously
 \begin{equation}
-\Ebb\biggl[\bigl(f(X') - f(X)\bigr)^2K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr] = \int_{\Xset} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx, \label{pf:first_order_graph_sobolev_seminorm_2}
-\end{equation}
-so that it remains now to bound the integral. Taking a first-order Taylor expansion of $f$, we get
+\Ebb\biggl[\bigl(f(X') - f(X)\bigr)^2K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr] \leq p_{\max}^2 \int_{\Xset} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx' \,dx, \label{pf:first_order_graph_sobolev_seminorm_2}
+\end{equation}
+so that it remains now to bound the double integral. Replacing difference by integrated derivative as in~\eqref{pf:first_order_graph_sobolev_seminorm_1.5}, we obtain
 \begin{align}
-\int_{\Xset} & \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \\
-& = \int_{\Xset} \int_{\Xset} \biggl[\int_{0}^{1} \nabla f\bigl(x + t(x' - x)\bigr)^T (x' - x)\,dt\biggr]^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \nonumber \\
-& \overset{(i)}{\leq} \int_{\Xset} \int_{\Xset} \int_{0}^{1} \biggl[\nabla f\bigl(x + t(x' - x)\bigr)^T (x' - x)\biggr]^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dt\,dx' \,dx \nonumber \\
-& \overset{(ii)}{=} r^{d + 2} \int_{\Xset} \int_{\mc{Z}_r(x)} \int_{0}^{1} \biggl[\nabla f\bigl(x + trz\bigr)^T z\biggr]^2 K\bigl(\|z\|\bigr) p(zr + x) p(x) \,dt\,dz \,dx \\
-&  \overset{(iii)}{\leq} r^{d + 2} \int_{\Xset} \int_{\Reals^d} \int_{0}^{1} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) p\bigl(\wt{x} + [1-t]rz\bigr) p\bigl(\wt{x} + trz\bigr) \,dt \,dz \,d\wt{x} \nonumber \\
-& \overset{(iv)}\leq \biggl(1 + \frac{L_pr}{p_{\min}}\biggr)^2 r^{d + 2} \int_{\Xset} \int_{\Reals^d} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) p(\wt{x})^2 \,dz \,d\wt{x} \label{pf:first_order_graph_sobolev_seminorm_3}
+\int_{\Xset} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx' \,dx & = \int_{\Xset} \int_{\Xset} \biggl[\int_{0}^{1} \nabla f\bigl(x + t(x' - x)\bigr)^{\top} (x' - x)\,dt\biggr]^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx' \,dx \nonumber \\
+& \overset{(i)}{\leq} \int_{\Xset} \int_{\Xset} \int_{0}^{1} \biggl[\nabla f\bigl(x + t(x' - x)\bigr)^{\top} (x' - x)\biggr]^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dt\,dx' \,dx \nonumber \\
+& \overset{(ii)}{\leq} r^{d + 2} \int_{\Xset} \int_{B(0,1)} \int_{0}^{1} \biggl[\nabla f\bigl(x + trz\bigr)^{\top} z\biggr]^2 K\bigl(\|z\|\bigr) \,dt\,dz \,dx \nonumber \\
+&  \overset{(iii)}{\leq} r^{d + 2} \int_{\Omega} \int_{B(0,1)} \int_{0}^{1} \Bigl[\nabla f\bigl(\wt{x}\bigr)^{\top} z\Bigr]^2 K\bigl(\|z\|\bigr) \,dt \,dz \,d\wt{x} \label{pf:first_order_graph_sobolev_seminorm_3}
 \end{align}
-where $(i)$ follows by Jensen's inequality, $(ii)$ follows by substituting $z = (x' - x)/r$ with $\mc{Z}_r(x) := \{z: zr + x \in \Xset\}$ $(iii)$ by substituting $\wt{x} = x + trz$ and noting that
-\begin{equation*}
-x \in \Xset ~~\textrm{and}~~ z \in \mc{Z}_r(x) \Longrightarrow x + trz \in \Xset ~~\textrm{and}~~ z \in \Reals^d,
-\end{equation*}
-and $(iv)$ by the Lipschitz property of $p$.
-
-Now, writing $\bigl(\nabla f(\wt{x}) ^T z\bigr)^2 = \bigl(\sum_{i = 1}^{d} z_{i} f^{(e_i)}(x) \bigr)^2$, expanding the square and integrating, we have that for any $\wt{x} \in \Xset$,
-\begin{align*}
-\int_{\Reals^d} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) \,dz & = \sum_{i,j = 1}^{d} f^{(e_i)}(\wt{x}) f^{(e_j)}(\wt{x}) \int_{\Rd} z_iz_jK(\|z\|) \,dz \\
-& = \sum_{i = 1}^{d} \bigl(f^{(e_i)}(\wt{x})\bigr)^2 \int_{\Rd} z_i^2 K\bigl(\|z\|\bigr) \,dz \\
-& = \sigma_K \|\nabla f(\wt{x})\|^2
-\end{align*}
-and by plugging back into~\eqref{pf:first_order_graph_sobolev_seminorm_3}, we obtain the claimed result.
+where $(i)$ follows by Jensen's inequality, $(ii)$ follows by substituting $z = (x' - x)/r$ and~\ref{asmp:kernel}, and $(iii)$ by exchanging integrals, substituting $\wt{x} = x + trz$, and noting that $x \in \mc{X}$ implies that $\wt{x} \in \Omega$.
+
+Now, writing $\bigl(\nabla f(\wt{x}) ^{\top} z\bigr)^2 = \bigl(\sum_{i = 1}^{d} z_{i} f^{(e_i)}(x) \bigr)^2$, expanding the square and integrating, we have that for any $\wt{x} \in \Xset$,
+\begin{align*}
+\int_{B(0,1)} \Bigl[\nabla f\bigl(\wt{x}\bigr)^{\top} z\Bigr]^2 K\bigl(\|z\|\bigr) \,dz & = \sum_{i,j = 1}^{d} f^{(e_i)}(\wt{x}) f^{(e_j)}(\wt{x}) \int_{\Rd} z_iz_jK(\|z\|) \,dz \\
+& = \sum_{i = 1}^{d} \bigl(f^{(e_i)}(\wt{x})\bigr)^2 \int_{B(0,1)} z_i^2 K\bigl(\|z\|\bigr) \,dz \\
+& = \sigma_K \|\nabla f(\wt{x})\|^2,
+\end{align*}
+where the last equality follows from the rotational symmetry of $K(\|z\|)$. Plugging back into~\eqref{pf:first_order_graph_sobolev_seminorm_3}, we obtain
+\begin{equation*}
+\int_{\Xset} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx' \,dx \leq r^{d + 2} \sigma_K \|f\|_{H^1(\Omega)}^2 \leq C_5 r^{d + 2} \sigma_K \|f\|_{H^1(\mc{X})}^2,
+\end{equation*}
+proving the claim of Lemma~\ref{lem:graph_sobolev_seminorm} upon taking $C_2 := C_5 \sigma_K p_{\max}^2$ in the statement of the lemma.
 
@@
 Suppose $f$ satisfies $|f(x') - f(x)| \leq M \|x - x\|$ for all $x,x' \in \mc{X}$.
-Then we can strengthen the high probability bound in Lemma~\ref{lem:graph_sobolev_seminorm} from $1 - \delta$ to $1 - \delta^2/n$, at the cost of only a constant factor in the upper bound on $f^T \Lap_{n,r} f$.
+Then we can strengthen the high probability bound in Lemma~\ref{lem:graph_sobolev_seminorm} from $1 - \delta$ to $1 - \delta^2/n$, at the cost of only a constant factor in the upper bound on $f^{\top} \Lap_{n,r} f$.
 \begin{proposition}
@@
 	\begin{equation*}
-	f^T \Lap_{n,r} f \leq \biggl(\frac{1}{\delta} + (1 + L_p)^2 \sigma_K\biggr) n^2 r^{d + 2} M^2
+	f^{\top} \Lap_{n,r} f \leq \biggl(\frac{1}{\delta} + C_2\biggr) n^2 r^{d + 2} M^2
 	\end{equation*}
@@
 \paragraph{Proof of Proposition~\ref{prop:graph_sobolev_seminorm_lipschitz}.}
-We will prove Proposition~\ref{prop:graph_sobolev_seminorm_lipschitz} using Chebyshev's inequality, so the key step is to upper bound the variance of $f^{T}\Lap_{n,r}f$.
-Putting $\varDelta_{ij} := K(\|X_i - X_j\|/r) \cdot (f(X_i) - f(X_j))^2$, we can write the variance of $f^T \Lap_{n,r} f$ as a sum of covariances,
-\begin{equation*}
-\Var\bigl[f^T \Lap_{n,r} f\bigr] = \frac{1}{4} \sum_{i, j = 1}^{n} \sum_{\ell,m = 1}^{n} \Cov\bigl[\varDelta_{ij},\varDelta_{\ell m}\bigr].
+We will prove Proposition~\ref{prop:graph_sobolev_seminorm_lipschitz} using Chebyshev's inequality, so the key step is to upper bound the variance of $f^{\top}\Lap_{n,r}f$.
+Putting $\varDelta_{ij} := K(\|X_i - X_j\|/r) \cdot (f(X_i) - f(X_j))^2$, we can write the variance of $f^{\top} \Lap_{n,r} f$ as a sum of covariances,
+\begin{equation*}
+\Var\bigl[f^{\top} \Lap_{n,r} f\bigr] = \frac{1}{4} \sum_{i, j = 1}^{n} \sum_{\ell,m = 1}^{n} \Cov\bigl[\varDelta_{ij},\varDelta_{\ell m}\bigr].
 \end{equation*}
@@
 	\item[$\bigl|I\bigr| = 4$.] In this case $\varDelta_{ij}$ and $\varDelta_{\ell m}$ are independent, and $\Cov\bigl[\varDelta_{ij},\varDelta_{\ell m}\bigr] = 0$.
-	\item[$\bigl|I\bigr| = 3$.] Taking $i = \ell$ without loss of generality, we have
+	\item[$\bigl|I\bigr| = 3$.] Taking $i = \ell$ without loss of generality, and noting that the expectation of $\Delta_{ij}$ and $\Delta_{im}$ is non-negative, we have
 	\begin{align*}
@@
 \begin{equation*}
-\Var\bigl[f^T \Lap_{n,r} f\bigr] \leq n^3 p_{\max}^3 M^4 r^{4 + 2d} + n^2 p_{\max}^2 M^4 r^{4 + d} K(0) \leq Cn^3r^{4 + 2d},
+\Var\bigl[f^{\top} \Lap_{n,r} f\bigr] \leq n^3 p_{\max}^3 M^4 r^{4 + 2d} + n^2 p_{\max}^2 M^4 r^{4 + d} K(0) \leq C M^4 n^3r^{4 + 2d},
 \end{equation*}
@@
 \begin{equation*}
-\Pbb\biggl(\Bigl|f^T \Lap_{n,r} f - \Ebb\bigl[f^T \Lap_{n,r} f\bigr]\Bigr| \geq \frac{M^2}{\delta}n^2 r^{d + 2}\biggr) \leq C\frac{\delta^2}{n}
-\end{equation*}
-and since we have already upper bounded $\Ebb\bigl[f^T \Lap_{n,r} f\bigr] \leq (1 + L_p)^2 \sigma_K M^2 n^2 r^{d + 2}$, the proposition follows.
+\Pbb\biggl(\Bigl|f^{\top} \Lap_{n,r} f - \Ebb\bigl[f^{\top} \Lap_{n,r} f\bigr]\Bigr| \geq \frac{M^2}{\delta}n^2 r^{d + 2}\biggr) \leq C\frac{\delta^2}{n}
+\end{equation*}
+and since we have already upper bounded $\Ebb\bigl[f^{\top} \Lap_{n,r} f\bigr] \leq C_2 M^2 n^2 r^{d + 2}$, the proposition follows.
 
@@
 \label{sec:graph_eigenvalues}
-In this section, we prove Lemma~\ref{lem:neighborhood_eigenvalue}, following the lead of~\citep{burago2014,trillos2019,calder2019}, who establish similar results with respect to a manifold without boundary.
-The bulk of our technical effort will stem from proving Theorem~\ref{thm:neighborhood_eigenvalue}, which gives estimates on eigenvalues of the graph Laplacian in terms of eigenvalues of the weighted Laplace-Beltrami operator $\Delta_P$.
+In this section, we prove Lemma~\ref{lem:neighborhood_eigenvalue}, following the lead of~\citet{burago2014,trillos2019,calder2019}, who establish similar results with respect to a manifold without boundary.
+To prove this lemma, in  Theorem~\ref{thm:neighborhood_eigenvalue} we give estimates on the difference between eigenvalues of the graph Laplacian $L_{n,r}$ and eigenvalues of the weighted Laplace-Beltrami operator $\Delta_P$.
 We recall $\Delta_P$ is defined as
 \begin{equation*}
-\Delta_Pf(x) := -\frac{1}{p} \dive\bigl(p^2\nabla f(x)\bigr)
-\end{equation*}
-To avoid confusion, in this section we write $\lambda_k(G_{n,r})$ for the $k$th smallest eigenvalue of the graph Laplacian matrix $\Lap_{n,r}$ and $\lambda_k(\Delta_P)$ for the $k$th smallest eigenvalue of $\Delta_P$ \footnote{Under the assumptions~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density}, the operator $\Delta_p$ has a discrete spectrum; see \cite{garciatrillos18} for more details.}.
-Some other notation: throughout this section, we will write $A, A_0, A_1,\ldots$ and $a,a_0,a_1,\ldots$ for constants which may depend on $\Xset$, $d$, $K$, and $p$ but do not depend on $n$.
-We let $L_K$ denote the Lipschitz constant of the kernel $K$, and we let $\theta$ and $\wt{\delta}$ denote (small) positive numbers, which always satisfy the following restrictions.
-\begin{enumerate}[label=(S\arabic*)]
-	\item 
+\Delta_Pf(x) := -\frac{1}{p(x)} \dive\bigl(p^2\nabla f\bigr)(x)
+\end{equation*}
+To avoid confusion, in this section we write $\lambda_k(G_{n,r})$ for the $k$th smallest eigenvalue of the graph Laplacian matrix $\Lap_{n,r}$ and $\lambda_k(\Delta_P)$ for the $k$th smallest eigenvalue of $\Delta_P$ \footnote{Under the assumptions~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density}, the operator $\Delta_P$ has a discrete spectrum; see \citet{garciatrillos18} for more details.}.
+Some other notation: throughout this section, we will write $A, A_0, A_1,\ldots$ and $a,a_0,a_1,\ldots$ for constants which may depend on $\Xset$, $d$, $K$, and $p$, but do not depend on $n$; we keep track of all such constants explicitly in our proofs.
+We let $L_K$ denote the Lipschitz constant of the kernel $K$.
+Finally, for notational ease we set $\theta$ and $\wt{\delta}$ to be the following (small) positive constants.
+\begin{equation}
 \label{asmp:smallness}
-	\begin{equation*}
-	\theta + \wt{\delta} \leq \frac{1}{4\max\{A_1,A\}}~~\textrm{and}~~n^{-1/d} \leq \wt{\delta} \leq \min\biggl\{\frac{1}{2^{d + 3}A_0}, \frac{1}{A_3}, \frac{K(1)}{8L_KA_0}\biggr\}r
-	\end{equation*}
-\end{enumerate}
+\wt{\delta} := \max\Biggl\{n^{-1/d}, \min\biggl\{\frac{1}{2^{d + 3}A_0}, \frac{1}{A_3}, \frac{K(1)}{8L_KA_0}, \frac{1}{8\max\{A_1,A\}c_0}\biggr\}r\Biggr\},~~\textrm{and}~~
+\theta := \frac{1}{8\max\{A_1,A\}}.
+\end{equation} 
+We note that each of $\wt{\delta}, \theta$ and $\wt{\delta}/r$ are of at most constant order.
 
@@
 	\label{thm:neighborhood_eigenvalue}
-	For any $\ell \in \mathbb{N}$ and $\theta$ and $\wt{\delta}$ such that
+	For any $\ell \in \mathbb{N}$ such that
 	\begin{equation}
@@
 	\end{equation}
-	with probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{m})$, it holds that
+	with probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{d})$, it holds that
 	\begin{equation}
@@
 \end{theorem}
-We will return to proving Theorem~\ref{thm:neighborhood_eigenvalue}, but first we show that under our regularity conditions on $p$ and $\Xset$, it implies Lemma~\ref{lem:neighborhood_eigenvalue}.
+We note that Theorem~\ref{thm:neighborhood_eigenvalue} is not a strong enough result to imply that $\lambda_{k}(G_{n,r})$ is a consistent estimate of $\lambda_k(\Delta_P)$, i.e.
+it does not imply that $|(nr^{d + 2})^{-1}\lambda_{k}(G_{n,r}) - \lambda_k(\Delta_P)| \to 0$ as $n \to \infty, r \to 0$.
+This is contrast to the results of~\citet{burago2014,trillos2019,calder2019}, whose results do imply consistency when the domain is assumed to be a manifold without boundary.
+While a more subtle analysis might imply consistency in our setting, the conclusion of Theorem~\ref{thm:neighborhood_eigenvalue}---that $\lambda_k(G_{n,r})/(nr^{d + 2}\lambda_k(\Delta_P))$ is bounded above and below by constants---suffices for our purposes.
+
+The bulk of this section is devoted to the proof of Theorem~\ref{thm:neighborhood_eigenvalue}.
+First, however, we show that our regularity under our conditions on $p$ and $\Xset$, Lemma~\ref{lem:neighborhood_eigenvalue} is a simple consequence of Theorem~\ref{thm:neighborhood_eigenvalue}.
 The link between the two is Weyl's Law.
@@
 \end{proposition}
-See Lemma 28 of~\citep{dunlop2020} for a proof that~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density} imply Weyl's Law.
+See Lemma 28 of~\citet{dunlop2020} for a proof that~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density} imply Weyl's Law.
 \paragraph{Proof of Lemma~\ref{lem:neighborhood_eigenvalue}.}
@@
 \begin{equation*}
-\ell_{\star} = \floor{\biggl(\frac{\bigl(1/(2A) - (\theta + \wt{\delta})\bigr)}{rA_2^{1/2}}\biggr)^d},
-\end{equation*}
-and note that the condition $r < c_0$ guarantees $\ell_{\star} \geq 2$ when $c_0$ is chosen to be a sufficiently small constant. By Proposition~\ref{prop:weyl}, we have that
-\begin{align*}
-\sqrt{\lambda_{\ell_{\star}}(\Delta_P)} \leq \frac{1}{r}\biggl(\frac{1}{2A} - (\theta + \wt{\delta})\biggr) 
-\end{align*}
-so that condition~\eqref{eqn:neighborhood_eigenvalue_1} is satisfied.
-We may therefore assume the conclusions of Theorem~\ref{thm:neighborhood_eigenvalue}.
-Along with~\eqref{eqn:weyls_law}, this implies the following bounds on the graph Laplacian eigenvalues:
+\ell_{\star} = \floor{\biggl(\frac{\bigl(1/(2A) - (\theta + \wt{\delta})\bigr)}{rA_2^{1/2}}\biggr)^d}.
+\end{equation*}
+Let us verify that $\lambda_{\ell_{\star}}(\Delta_P)$ satisfies the condition~\ref{eqn:neighborhood_eigenvalue_1} of Theorem~\ref{thm:neighborhood_eigenvalue}.
+Setting $c_0 := 1/(2^{1/d}4A_2^{1/2})$, the assumed upper bound on the radius $r < c_0$ in \ref{asmp:radius1} guarantees that $\ell_{\star} \geq 2$.
+Therefore, by Proposition~\ref{prop:weyl} we have that
+\begin{align*}
+\sqrt{\lambda_{\ell_{\star}}(\Delta_P)} \leq A_2^{1/2}\ell_{\star}^{1/d} \leq  \frac{1}{r}\biggl(\frac{1}{2A} - (\theta + \wt{\delta})\biggr).
+\end{align*}
+Rearranging the above inequality shows that condition~\eqref{eqn:neighborhood_eigenvalue_1} is satisfied.
+
+It is therefore the case that the inequalities in~\eqref{eqn:eigenvalue_bound} hold with probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{m})$.
+Together, \eqref{eqn:eigenvalue_bound} and \eqref{eqn:weyls_law} imply the following bounds on the graph Laplacian eigenvalues:
 \begin{equation*}
@@
 On the one hand, since the eigenvalues are sorted in ascending order, we can use the lower bound on $\lambda_{\ell_{\star}}(G_{n,r})$ that we have just derived:
-\begin{align*}
-\lambda_k(G_{n,r}) & \geq \lambda_{\ell_{\star}}(G_{n,r}) \\
-& \geq \frac{a_2}{A}nr^{d + 2}\ell_{\star}^{2/d} \\
-& = \frac{a_2}{A}nr^{d + 2} \floor{\biggl(\frac{\bigl(1/(2A) - (\theta + \wt{\delta})\bigr)}{rA_2^{1/2}}\biggr)^d}^{2/d} \\
-& \geq \frac{a_2}{32A^3 A_2} nr^{d}
-\end{align*}
-with the last line following since $\theta + \wt{\delta} \leq \frac{1}{4A}$.
+\begin{equation*}
+\lambda_k(G_{n,r}) \geq \lambda_{\ell_{\star}}(G_{n,r}) \geq \frac{a_2}{A}nr^{d + 2}\ell_{\star}^{2/d} \geq \frac{a_2}{64A^3 A_2} nr^{d}.
+\end{equation*}
 On the other hand, for any graph $G$ the maximum eigenvalue of the Laplacian is upper bounded by twice the maximum degree \citep{chung97}.
-Writing $D_{\max}(G_{n,r})$ for the maximum degree of $G_{n,r}$, we show in Lemma~\ref{lem:max_degree} that
+Writing $D_{\max}(G_{n,r})$ for the maximum degree of $G_{n,r}$, it is thus a consequence of Lemma~\ref{lem:max_degree} that
 \begin{equation*}
@@
 with probability at least $1 - 2n\exp\Bigl(-nr^dp_{\min}/(3K(0)^2)\Bigr)$.
-Setting $\theta$ and $\wt{\delta}$ as large as possible while ensuring that~\ref{asmp:smallness} and~\eqref{eqn:neighborhood_eigenvalue_1} are satisfied---so that $\theta \asymp 1$ and $\wt{\delta} \asymp r$---we have established that each of the inequalities in Lemma~\ref{lem:neighborhood_eigenvalue} hold with probability at least $1 - Cn\exp(-cnr^d)$.
+In sum, we have shown that with probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{d}) - 2n\exp\Bigl(-nr^dp_{\min}/(3K(0)^2)\Bigr)$,
+\begin{equation*}
+\min\biggl\{\frac{a_2}{A}nr^{d + 2}k^{2/d}, \frac{a_2}{A^3 64 A_3} nr^d\biggr\} \leq \lambda_k(G_{n,r}) \leq \min\biggl\{\frac{A_2}{a}nr^{2 + d}k^{2/d}, 4p_{\max}nr^d\biggr\}~~\textrm{for all $2 \leq k \leq n$.}
+\end{equation*}
+Lemma~\ref{lem:neighborhood_eigenvalue} then follows upon setting
+\begin{align*}
+C_1 & := \max\{2A_0,4\},~~ && c_1 := \min\Biggl\{\frac{p_{\min}}{3K(0)^2}, \frac{\theta^2\wt{\delta}}{r}\Biggr\} \\
+C_3 & := \max\biggl\{\frac{A_2}{a}, 4p_{\max}\biggr\},~~ && c_3 := \min\biggl\{\frac{a_2}{A}, \frac{a_2}{A^3 64 A_3} \biggr\}.
+\end{align*}
+in the statement of that Lemma.
 
@@
 
-In this section we prove Theorem~\ref{thm:neighborhood_eigenvalue}, following closely the approach of \citep{burago2014,trillos2019,calder2019}.
+In this section we prove Theorem~\ref{thm:neighborhood_eigenvalue}, following closely the approach of \citet{burago2014,trillos2019,calder2019}.
 As in these works, we relate $\lambda_k(\Delta_P)$ and $\lambda_k(G_{n,r})$ by means of the Dirichlet energies
 \begin{equation*}
-b_r(u) := \frac{1}{n^2 r^{d+ 2}}u^T \Lap_{n,r} u 
+b_r(u) := \frac{1}{n^2 r^{d+ 2}}u^{\top} \Lap_{n,r} u 
 \end{equation*}
@@
 \end{equation*}
-We recall that for a function $u: \{X_1,\ldots,X_n\} \to \Reals$, the empirical norm is defined as $\|u\|_n^2 := \frac{1}{n} \sum_{i = 1}^{n} (u(X_i))^2$, and we let the class $\Leb^2(P_n)$ consist of those $u \in \Reals^n$ for which $\|u\|_{n} < \infty$.
-For a function $f: \Xset \to \Reals$, the continuum limit of $\|f\|_n^2$ is
-\begin{equation*}
-\|f\|_{P}^2 := \int_{\Xset} \bigl|f(x)\bigr|^2 p(x) \,dx.
-\end{equation*}
-
-Let us pause briefly to motivate the relevance of $b_r(u)$ and $D_2(f)$. Suppose one could show the following two results: 
+Let us pause briefly to motivate the relevance of $b_r(u)$ and $D_2(f)$.
+In the following discussion, recall that for a function $u: \{X_1,\ldots,X_n\} \to \Reals$, the empirical norm is defined as $\|u\|_n^2 := \frac{1}{n} \sum_{i = 1}^{n} (u(X_i))^2$, and the class $\Leb^2(P_n)$ consists of those $u \in \Reals^n$ for which $\|u\|_{n} < \infty$.
+Similarly, for a function $f: \Xset \to \Reals$, the $L^2(P)$ norm of $f$ is
+\begin{equation*}
+\|f\|_{P}^2 := \int_{\Xset} \bigl|f(x)\bigr|^2 p(x) \,dx,
+\end{equation*} 
+and the class $\Leb^2(P)$ consists of those $f$ for which $\|f\|_P < \infty$. Now, suppose one could show the following two results: 
 \begin{enumerate}[(1)]
@@
 We will momentarily define particular maps $\wt{\mc{I}}$ and $\wt{\mc{P}}$, and establish that they satisfy both (1) and (2).
-In order to define these maps, we must first introduce a particular probability measure $\wt{P}_n$ that with high probability is close in transportation distance to both $P_n$ and $P$.
-This estimate on the transportation distance---which we now give--- will be the workhorse that allows us to relate $b_r$ to $D_2$, and $\|\cdot\|_n$ to $\|\cdot\|_P$.
+In order to define these maps, we must first introduce a particular probability measure $\wt{P}_n$ that, with high probability, is close in transportation distance to both $P_n$ and $P$.
+This estimate on the transportation distance---which we now give---will be the workhorse that allows us to relate $b_r$ to $D_2$, and $\|\cdot\|_n$ to $\|\cdot\|_P$.
 
@@
 
-\cite{calder2019} take $\Xset$ to be a smooth submanifold of $\Rd$ without boundary, i.e.
+\citet{calder2019} take $\Xset$ to be a smooth submanifold of $\Rd$ without boundary, i.e.
 they assume $\Xset$ satisfies~\ref{asmp:domain_manifold}.
-They exhibit an absolutely continuous measure $\wt{P}_n$ with density $\wt{p}_n$ which with high probability is close to $P_n$ in transportation distance, for which $\|p - \wt{p}_n\|_{\Leb^\infty}$ is also small.
+In this setting, they exhibit an absolutely continuous measure $\wt{P}_n$ with density $\wt{p}_n$ that with high probability is close to $P_n$ in transportation distance, and for which $\|p - \wt{p}_n\|_{\Leb^\infty}$ is also small.
 In Proposition~\ref{prop:optimal_transport}, we adapt this result to the setting of full-dimensional manifolds with boundary.
@@
 Additionally we denote by $\wt{T}_n$ an \emph{optimal transport map} between $\wt{P}_n$ and $P_n$, meaning a transportation map which achieves the infimum in~\eqref{eqn:optimal_transport}.
-Finally, we write $U_1,\ldots,U_n$ for the preimages of $X_1,\ldots,X_n$ under $\wt{T}_n$.
+Finally, we write $U_1,\ldots,U_n$ for the preimages of $X_1,\ldots,X_n$ under $\wt{T}_n$, meaning $U_i = \wt{T}_n^{-1}(X_i)$.
 
 \paragraph{Interpolation and discretization maps.}
-The discretization map  $\wt{\mathcal{P}}: \Leb^2(\Xset) \to \Leb^2(P_n)$ is given by
+The discretization map  $\wt{\mathcal{P}}: \Leb^2(\Xset) \to \Leb^2(P_n)$ is given by averaging over the cells $U_1,\ldots,U_n$, 
 \begin{equation*}
@@
 \end{equation*} 
-and $\Lambda_{r - 2A_0\wt{\delta}}$ is a smoothening operator. To be precise, for any $h > 0$,
+and $\Lambda_{r - 2A_0\wt{\delta}}$ is a kernel smoothing operator, defined with respect to a carefully chosen kernel $\psi$. To be precise, for any $h > 0$,
 \begin{equation*}
@@
 	\begin{enumerate}[(1)]
-		\item For any $u \in \Leb^2(P_n)$,
-		\begin{equation*}
+		\item For every $u \in \Leb^2(P_n)$,
+		\begin{equation}
+		\label{eqn:dirichlet_energies_1}
 		\sigma_{K} D_2(\wt{\mc{I}}u) \leq A_8 \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) b_r(u)
-		\end{equation*}
-		\item For any $f \in \Leb^2(\Xset)$,
-		\begin{equation*}
-		b_r(\wt{\mc{P}}f) \leq \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + \frac{L_p(r + 2A_0\wt{\delta})}{p_{\min}}\Bigr)^2 \cdot \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr) \cdot \sigma_{K} D_2(f)
-		\end{equation*}
+		\end{equation}
+		\item For every $f \in \Leb^2(\Xset)$,
+		\begin{equation}
+		\label{eqn:dirichlet_energies_2}
+		b_r(\wt{\mc{P}}f) \leq \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr) \cdot \Bigl(\frac{C_5 p_{\max}^2}{p_{\min}^2}\Bigr) \cdot \sigma_{K} D_2(f)
+		\end{equation}
 	\end{enumerate}
@@
 		\item For every $f \in \Leb^2(\Xset)$,
-		\begin{equation*}
-		\Bigl|\|f\|_{P}^2 - \|\wt{P}f\|_{n}^2\Bigr| \leq A_5 r \|f\|_{P} \sqrt{D_2(f)} + A_1\bigl(\theta + \wt{\delta}\bigr)\|f\|_{P}^2
-		\end{equation*}
+		\begin{equation}
+		\label{eqn:isometry_1}
+		\Bigl|\|f\|_{P}^2 - \|\wt{\mc{P}}f\|_{n}^2\Bigr| \leq A_5 r \|f\|_{P} \sqrt{D_2(f)} + A_1\bigl(\theta + \wt{\delta}\bigr)\|f\|_{P}^2
+		\end{equation}
 		\item For every $u \in \Leb^2(P_n)$,
-		\begin{equation*}
+		\begin{equation}
+		\label{eqn:isometry_2}
 		\Bigl|\|\wt{\mc{I}}u\|_{P}^2 - \|u\|_{n}^2\Bigr| \leq A_6 r \|u\|_{n} \sqrt{b_r(u)} + A_7\bigl(\theta + \wt{\delta}\bigr) \|u\|_{n}^2
-		\end{equation*}
+		\end{equation}
 	\end{enumerate}
@@
 \paragraph{Proof of Theorem~\ref{thm:neighborhood_eigenvalue}.}
+Throughout this proof, we assume that inequalities~\eqref{eqn:dirichlet_energies_1}-\eqref{eqn:isometry_2} are satisfied. We take $A$ and $a$ to be positive constants such that
+\begin{align*}
+\frac{1}{a} & \geq 2\Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)\Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr)\Bigl(\frac{C_5 p_{\max}^2}{p_{\min}^2}\Bigr) ,~~\textrm{and}~~A \geq \max\Biggl\{A_1,A_5, \frac{1}{\sqrt{a}}A_6, A_7, 2A_8\biggl(1 + A_1(\theta + \wt{\delta})\biggr) \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) \Biggr\}.
+\end{align*}
+
 Let $k$ be any number in $1,\ldots,\ell$.
-We start with the upper bound, proceeding exactly as in Proposition 4.4 of \citep{burago2014}.
-Let $f_1,\ldots,f_{k}$ denote the first $k$ eigenfunctions of $\Delta_P$ and set $W := \mathrm{span}(f_1,\ldots,f_k)$, so that by the Courant-Fischer principle $D_2(f) \leq \lambda_k(\Delta_P) \|f\|_{P}^2$ for every $f \in W$.
+We start with the upper bound in~\eqref{eqn:eigenvalue_bound}, proceeding as in Proposition 4.4 of \citet{burago2014}.
+Let $f_1,\ldots,f_{k}$ denote the first $k$ eigenfunctions of $\Delta_P$ and set $W := \mathrm{span}\{f_1,\ldots,f_k\}$, so that by the Courant-Fischer principle $D_2(f) \leq \lambda_k(\Delta_P) \|f\|_{P}^2$ for every $f \in W$.
 As a result, by Part (1) of Proposition~\ref{prop:isometry} we have that for any $f \in W$,
@@
 \end{equation*}
-where the second inequality follows by assumption~\eqref{eqn:neighborhood_eigenvalue_1}, upon appropriate choice of $A$.
+where the second inequality follows by assumption~\eqref{eqn:neighborhood_eigenvalue_1}.
 
@@
 & = \max_{\substack{f \in W \\ f \neq 0} } \frac{b_r(\wt{\mc{P}}f)}{\bigl\|\wt{\mc{P}}f\bigr\|_{n}^2} \\
-& \leq 2\Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + \frac{L_p(r + 2A_0\wt{\delta})}{p_{\min}}\Bigr)^2 \cdot \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr)\sigma_K \lambda_{k}(\Delta_P),
-\end{align*}
-establishing the lower bound in~\eqref{eqn:eigenvalue_bound} upon proper choice of $a$.
+& \leq 2\Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr) \cdot \Bigl(\frac{C_5 p_{\max}^2}{p_{\min}^2}\Bigr) \sigma_K \lambda_{k}(\Delta_P),
+\end{align*}
+establishing the lower bound in~\eqref{eqn:eigenvalue_bound}.
 
 The upper bound follows from essentially parallel reasoning.
-Recalling that $v_1,\ldots,v_k$ denote the first $k$ eigenvectors of $\Lap_{n,r}$, set $U := \mathrm{span}(v_1,\ldots,v_k)$, so that $nr^{d + 2} b_r(u) \leq \lambda_k(G_{n,r}) \|u\|_n^2$.
-By Proposition~\ref{prop:isometry}, Part (2), for every $u \in U$ it holds that
+Recalling that $v_1,\ldots,v_k$ denote the first $k$ eigenvectors of $\Lap_{n,r}$, set $U := \mathrm{span}\{v_1,\ldots,v_k\}$, so that $nr^{d + 2} b_r(u) \leq \lambda_k(G_{n,r}) \|u\|_n^2$.
+By Proposition~\ref{prop:isometry}, Part (2), we have that for every $u \in U$,
 \begin{align*}
@@
 \end{align*}
-where the second to last inequality follows from the lower bound $a \lambda_k(G_{n,r}) \leq nr^{d + 2}\lambda_k(\Delta_P)$ that we just derived, and the last inequality from assumption~\eqref{eqn:neighborhood_eigenvalue_1}, upon appropriate choice of $A$.
+where the second to last inequality follows from the lower bound $a \lambda_k(G_{n,r}) \leq nr^{d + 2}\lambda_k(\Delta_P)$ that we just derived, and the last inequality from assumption~\eqref{eqn:neighborhood_eigenvalue_1}.
 
@@
 \end{align*}
-establishing the upper bound in~\eqref{eqn:eigenvalue_bound} when $A$ is properly chosen.
+establishing the upper bound in~\eqref{eqn:eigenvalue_bound}.
 
@@
 
-The proof of Proposition~\ref{prop:optimal_transport} is given in Subsection~\ref{subsec:proof_proposition_optimal_transport}.
-In Subsection~\ref{subsec:integrals}, we establish relationships between the (non-random) functionals $E_r(f)$ and $D_2(f)$, as well as providing estimates on some assorted integrals.
-In Subsection~\ref{subsec:random_functionals}, we establish relationships between the stochastic functionals $\wt{E}_r(f)$ and $E_r(f)$,  between $\wt{E}_r\bigl(\wt{\mc{I}}(u)\bigr)$ and $b_r\bigl(u\bigr)$, and between $\wt{E}_r\bigl(f\bigr)$ and $b_r\bigl(\wt{\mc{P}}f\bigr)$.
-Finally, in Subsection~\ref{subsec:proof_of_prop_dirichlet_energies_and_isometry} we use these various relationships to prove Propositions~\ref{prop:dirichlet_energies} and~\ref{prop:isometry}.
+The proof of Proposition~\ref{prop:optimal_transport} is given in Section~\ref{subsec:proof_proposition_optimal_transport}.
+In Section~\ref{subsec:integrals}, we establish relationships between the (non-random) functionals $E_r(f)$ and $D_2(f)$, as well as providing estimates on some assorted integrals.
+In Section~\ref{subsec:random_functionals}, we establish relationships between the stochastic functionals $\wt{E}_r(f)$ and $E_r(f)$,  between $\wt{E}_r\bigl(\wt{\mc{I}}(u)\bigr)$ and $b_r\bigl(u\bigr)$, and between $\wt{E}_r\bigl(f\bigr)$ and $b_r\bigl(\wt{\mc{P}}f\bigr)$.
+Finally, in Section~\ref{subsec:proof_of_prop_dirichlet_energies_and_isometry} we use these various relationships to prove Propositions~\ref{prop:dirichlet_energies} and~\ref{prop:isometry}.
 
@@
 	where $Q_k = \wt{\delta}\Bigl([k_1 - 1,k_1] \otimes \cdots \otimes [k_d - 1,k_d]\Bigr)$ and we assume without loss of generality that $\wt{\delta}^{-1} \in \mathbb{N}$.
-	\item If $\mc{\Xset}$ is an open, connected set with smooth boundary, then by Proposition 3.2 of \cite{trillos2015}, there exist a finite number $N(\Xset) \in \mathbb{N}$ of disjoint polytopes which cover $\Xset$.
+	\item If $\mc{\Xset}$ is an open, connected set with smooth boundary, then by Proposition 3.2 of \citet{trillos2015}, there exist a finite number $N(\Xset) \in \mathbb{N}$ of disjoint polytopes which cover $\Xset$.
 Moreover, letting $U_j$ denote the intersection of the $j$th of these polytopes with $\wb{\Xset}$, this proposition establishes that for each $j$ there exists a bi-Lipschitz homeomorphism $\Phi_j: U_j \to [0,1]^d$.
@@
 \end{equation}
-As a special case: when $\eta(x) = 1$, this implies $\vol\bigl(B(x,\varepsilon) \cap \Xset\bigr) \geq a_3 \nu_d \varepsilon^d$ for any $0 \varepsilon \leq c_0$.
+As a special case: when $\eta(x) = 1$, this implies $\vol\bigl(B(x,\varepsilon) \cap \Xset\bigr) \geq a_3 \nu_d \varepsilon^d$ for any $0 < \varepsilon \leq c_0$.
 
@@
 In Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, we establish the reverse inequality.
-\begin{lemma}[c.f. Lemma 9 of \cite{trillos2019}, Lemma 5.5 of \cite{burago2014}]
+\begin{lemma}[\textbf{c.f. Lemma 9 of \citet{trillos2019}, Lemma 5.5 of \citet{burago2014}}]
 	\label{lem:first_order_graph_sobolev_seminorm_expected_lb}
@@
 To prove Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, we require upper and lower bounds on $\tau_h(x)$, as well as an upper bound on the gradient of $\tau_h$.
-The lower bound here---$\tau_h(x) \geq 1/2$---is quite a bit a looser than what can be shown when $\Xset$ has no boundary.
+The lower bound here---$\tau_h(x) \geq a_3$---is quite a bit a looser than what can be shown when $\Xset$ has no boundary.
 The same is the case regarding the upper bound of the size of the gradient $\|\nabla \tau_h(x)\|$.
@@
 Finally, to prove part (2) of Proposition~\ref{prop:isometry}, we require Lemma~\ref{lem:smoothening_error}, which gives an estimate on the error $\Lambda_h f - f$ in $\|\cdot\|_P^2$ norm.
-\begin{lemma}[c.f Lemma 8 of \cite{trillos2019}, Lemma 5.4 of \cite{burago2014}]
+\begin{lemma}[\textbf{c.f Lemma 8 of \citet{trillos2019}, Lemma 5.4 of \citet{burago2014}}]
 	\label{lem:smoothening_error}
@@
 \begin{align*}
-\bigl(\nabla\eta_h(x',\cdot)\bigr)(x) & = \frac{1}{h^{d + 1}} \psi'\biggl(\frac{\|x'  - x\|}{h}\biggr) \frac{(x - x')}{\|x' - x\|} \\
-& = \frac{1}{\sigma_Kh^{d + 2}} K\biggl(\frac{\|x' - x\|}{h}\biggr) (x' - x),
+\bigl(\nabla\eta_h(x',\cdot)\bigr)(x) & = \frac{1}{h} \psi'\biggl(\frac{\|x'  - x\|}{h}\biggr) \frac{(x - x')}{\|x' - x\|} \\
+& = \frac{1}{\sigma_Kh^{2}} K\biggl(\frac{\|x' - x\|}{h}\biggr) (x' - x),
 \end{align*}
@@
 \begin{align*}
-\bigl\|J_1(x)\bigr\|^2 & = \frac{1}{\sigma_K^2 h^{4 + 2d}} \Biggl[\int_{\Xset} \bigl(f(x') - f(x)\bigr)K\biggl(\frac{\|x' - x\|}{h}\biggr)(x' - x)^Tw\,dx'\Biggr]^2 \\
-& \leq \frac{1}{\sigma_K^2 h^{4 + 2d}} \biggl[\int_{\Xset}K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl((x' - x)^T w\bigr)^2\,dx'\biggr] \biggl[\int_{\Xset}K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl(f(x') - f(x)\bigr)^2\,dx'\biggr].
+\bigl\|J_1(x)\bigr\|^2 & = \frac{1}{\sigma_K^2 h^{4 + 2d}} \Biggl[\int_{\Xset} \bigl(f(x') - f(x)\bigr)K\biggl(\frac{\|x' - x\|}{h}\biggr)(x' - x)^{\top}w\,dx'\Biggr]^2 \\
+& \leq \frac{1}{\sigma_K^2 h^{4 + 2d}} \biggl[\int_{\Xset}K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl((x' - x)^{\top} w\bigr)^2\,dx'\biggr] \biggl[\int_{\Xset}K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl(f(x') - f(x)\bigr)^2\,dx'\biggr].
 \end{align*}
@@
 \begin{align*}
-\int_{\Xset}K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl((x' - x)^T w\bigr)^2\,dx' & \leq \int_{\Rd}K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl((x' - x)^T w\bigr)^2\,dx' \\
-& = h^{d + 2} \int_{\Rd} K\bigl(\|z\|\bigr) \bigl(z^T w\bigr)^2 \,dz \\
-& = \sigma_K h^{d + 2},
+\int_{\Xset}K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl((x' - x)^{\top} w\bigr)^2\,dx' & \leq h^{d + 2} \int_{\Xset} K\bigl(\|z\|\bigr) \bigl(z^{\top} w\bigr)^2 \,dz \leq \sigma_K h^{d + 2},
 \end{align*}
@@
 & \leq \Bigl\|\nabla\biggl(\frac{1}{\tau_h}\biggr)(x)\Bigr\|^2\frac{\tau_h(x)}{\sigma_K h^d}\int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx' \\
-& \leq \frac{1}{da_3^3\sigma_K^2h^{2 + d}} \biggl[\int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr].
+& \leq \frac{1}{da_3^2\sigma_K^2h^{2 + d}} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx',
 \end{align*}
@@
 & \leq 2 \sigma_K \int_{\Xset} \Biggl(\frac{\bigl\|J_1(x)\bigr\|^2}{\tau_h^2(x)} + \bigl\|J_2(x)\bigr\|^2\Biggr) p^2(x) \,dx \\
-& \leq \biggl(\frac{1}{a_3^2} + \frac{1}{da_3^3\sigma_K}\biggr)\frac{2}{h^{d + 2}} \int_{\Xset} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl(f(x') - f(x)\bigr)^2 p^2(x) \,dx' \,dx \\
-& \leq 2 \Bigl(1 + \frac{L_ph}{p_{\min}}\Bigr) \biggl(\frac{1}{a_3^2} + \frac{1}{da_3^3\sigma_K}\biggr) E_h(f),
-\end{align*}
-and taking $A_8 := 2 \Bigl(1 + \frac{L_pc_0}{p_{\min}}\Bigr) \biggl(\frac{1}{a_3^2} + \frac{1}{da_3^3\sigma_K}\biggr)$ completes the proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}.
+& \leq \biggl(\frac{1}{a_3^2} + \frac{1}{da_3^2\sigma_K}\biggr)\frac{2}{h^{d + 2}} \int_{\Xset} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr)\bigl(f(x') - f(x)\bigr)^2 p^2(x) \,dx' \,dx \\
+& \leq 2 \Bigl(1 + \frac{L_ph}{p_{\min}}\Bigr) \biggl(\frac{1}{a_3^2} + \frac{1}{da_3^2\sigma_K}\biggr) E_h(f),
+\end{align*}
+and taking $A_8 := 2 \Bigl(1 + \frac{L_pc_0}{p_{\min}}\Bigr) \biggl(\frac{1}{a_3^2} + \frac{1}{da_3^2\sigma_K}\biggr)$ completes the proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}.
 
@@
 We will use Lemma~\ref{lem:poincare} in the proof of Proposition~\ref{prop:isometry}. 
-\begin{lemma}[c.f. Lemma 3.4 of \citep{burago2014}]
+\begin{lemma}[\textbf{c.f. Lemma 3.4 of \citet{burago2014}}]
 	\label{lem:poincare}
@@
 Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized} relates the graph Sobolev semi-norm $b_r(\wt{\mc{P}}f)$ to the non-local energy $\wt{E}_r(f)$. 
-\begin{lemma}[\textbf{c.f. Lemma 13 of \cite{trillos2019}, Lemma 4.3 of \cite{burago2014}}]
+\begin{lemma}[\textbf{c.f. Lemma 13 of \citet{trillos2019}, Lemma 4.3 of \citet{burago2014}}]
 	\label{lem:first_order_graph_sobolev_seminorm_discretized}
@@
 In Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}, we establish the reverse of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}. 
-\begin{lemma}[c.f. Lemma 14 of \cite{trillos2019}]
+\begin{lemma}[\textbf{c.f. Lemma 14 of \citet{trillos2019}}]
 	\label{lem:first_order_graph_sobolev_seminorm_discretized_lb}
@@
 \begin{equation*}
-\vol\bigl(V \cap \Xset\bigr) \geq \vol\bigl(B(x,r - 2A_0\wt{\delta})\bigr) \geq a_3 \nu_d (r - 2A_0\wt{\delta})^d \geq \frac{a_3 \nu_d}{2^d}r^d 
+\vol\bigl(V \cap \Xset\bigr) \geq \vol\bigl(B(x,r - 2A_0\wt{\delta}) \cap \mc{X}\bigr) \geq a_3 \nu_d (r - 2A_0\wt{\delta})^d \geq \frac{a_3 \nu_d}{2^d}r^d 
 \end{equation*}
@@
 we have that for any $x$ and $x'$ in $U$,
-\begin{align*}
-\bigl|f(x') - f(x)\big|^2 & \leq \frac{2}{\vol(V \cap \Xset)} \int_{V \cap \Xset} \bigl|f(x') - f(z)\big|^2 + \bigl|f(z) - f(x)\big|^2 \,dz \\
-& \leq \frac{2^{d + 1}}{a_3 \nu_d r^d} \int_{V \cap \Xset} \bigl|f(x') - f(z)\big|^2 + \bigl|f(z) - f(x)\big|^2 \,dz \\
-& \leq \frac{2^{d + 2}}{K(1) a_3 \nu_d r^d p_{\min}} \Bigl(F(x') + F(x)\bigr)
-\end{align*}
-where
-\begin{equation*}
-F(x) := \int_{\Xset} K\biggl(\frac{\|z - x\|}{r}\biggr) \bigl(f(z) - f(x)\bigr)^2 \wt{p}_n(x) \,dx.
-\end{equation*}
-Plugging this upper bound back into~\eqref{pf:poincare_1}, we have that
+\begin{align}
+\bigl|f(x') - f(x)\big|^2 & \leq \frac{2}{\vol(V \cap \Xset)} \int_{V \cap \Xset} \bigl|f(x') - f(z)\big|^2 + \bigl|f(z) - f(x)\big|^2 \,dz \nonumber \\
+& \leq \frac{2^{d + 1}}{a_3 \nu_d r^d} \int_{V \cap \Xset} \bigl|f(x') - f(z)\big|^2 + \bigl|f(z) - f(x)\big|^2 \,dz \nonumber \\
+& \leq \frac{2^{d + 2}}{K(1) a_3 \nu_d r^d p_{\min}} \Bigl(F(x') + F(x)\bigr) \label{pf:poincare_2}
+\end{align}
+where in the last inequality we set
+\begin{equation*}
+F(x) := \int_{\Xset} K\biggl(\frac{\|z - x\|}{r}\biggr) \bigl(f(z) - f(x)\bigr)^2 \wt{p}_n(x) \,dx,
+\end{equation*}
+and use the facts that $\wt{p}_n(x) \geq p_{\min}/2$, that $K(\|z  - x\|/r) \geq K(1)$ for all $z \in B(x,r)$.
+
+Plugging the upper bound~\eqref{pf:poincare_2} back into~\eqref{pf:poincare_1}, we have that
 \begin{align*}
@@
 \paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}.}
-Recalling that $\bigl(\wt{\mc{P}}f\bigr)(X_i) = n \cdot \int_{U_i} f(x) \wt{p}_n(x)$, by Jensen's inequality,
+Recalling that $\bigl(\wt{\mc{P}}f\bigr)(X_i) = n \cdot \int_{U_i} f(x) \wt{p}_n(x) \,dx$, by Jensen's inequality,
 \begin{equation*}
@@
 \end{equation*}
-Using this along with the Lipschitz property of $K$, we get
+As a result, the graph Dirichlet energy is upper bounded as follows:
 \begin{align*}
@@
 where the second inequality follows from the algebraic identities $(1 - t)^{-1} \leq (1 + 2t)$ for any $0 < t < 1/2$ and $(1 + s)(1 + t) < 1 + 2s + t$ for any $0 < t < 1$ and $s > 0$.
+The Lemma follows upon choosing $A_3 := \frac{8A_0L_K}{K(1)} + 2^{d + 3}$.
 
@@
 & \overset{(v)}{\leq} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr){E}_{r + 2A_0\wt{\delta}}(f) \\
-& \overset{(vi)}{\leq} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + \frac{L_p(r + 2A_0\wt{\delta})}{p_{\min}}\Bigr)^2 \cdot \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr) \cdot \sigma_{K} D_2(f)
+& \overset{(vi)}{\leq} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr) \cdot \Bigl(\frac{C_5 p_{\max}^2}{p_{\min}^2}\Bigr) \cdot \sigma_{K} D_2(f)
 \end{align*}
@@
 \begin{equation*}
-\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\wt{P}_n}^2 \leq A_3r^2 \wt{E}_r(f,\Xset) \leq  A_3 \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + \frac{L_p(r + 2A_0\wt{\delta})}{p_{\min}}\Bigr)^2r^2 D_2(f)
+\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\wt{P}_n}^2 \leq A_3r^2 \wt{E}_r(f,\Xset) \leq  A_3 \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(\frac{C_5 p_{\max}^2}{p_{\min}^2}\Bigr) \cdot \sigma_{K}r^2 D_2(f)
 \end{equation*}
@@
 \begin{equation*}
-A_5 := 2\sqrt{A_3} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \Bigl(1 + \frac{L_p(r + 2A_0\wt{\delta})}{p_{\min}}\Bigr).
+A_5 := 2\sqrt{A_3} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \Bigl(\frac{\sqrt{C_5} p_{\max}}{p_{\min}}\Bigr) \cdot \sqrt{\sigma_{K}}.
 \end{equation*}
@@
 	\label{lem:empirical_norm_sobolev}
-	Suppose $P$ satisfies~\ref{asmp:bounded_lipschitz_density}. If $f \in H^1(\Xset,M)$ is lower bounded in $\Leb^2(\Xset)$ norm,
+	Fix $\delta \in (0,1)$ Suppose $P$ satisfies~\ref{asmp:bounded_lipschitz_density}. If $f \in H^1(\Xset,M)$ is lower bounded in $\Leb^2(\Xset)$ norm,
 	\begin{equation}
 	\label{eqn:empirical_norm_sobolev_1}
-	\norm{f}_{\Leb^2(\Xset)} \geq 
-	\begin{cases*}
-	\frac{C_6 M}{\delta} \cdot \max\Bigl\{n^{-1/2},n^{-1/d}\Bigr\},& ~~\textrm{if $2 \neq d$} \\
-	\frac{C_6 M}{\delta} \cdot n^{-a/2},& ~~\textrm{if $2 = d$, for any $0 < a < 1$}
-	\end{cases*}
-	\end{equation}
-	for some $\delta \leq 1$, then
+	\norm{f}_{\Leb^2(\Xset)} \geq \frac{C_6 M}{\delta} \cdot \max\Bigl\{n^{-1/2},n^{-1/d}\Bigr\}
+	\end{equation}
+	Then with probability at least $1 - 5 \delta$,
 	\begin{equation}
 	\label{eqn:empirical_norm_sobolev}
-	\norm{f}_n^2 \geq \delta \cdot \Ebb\Bigl[\norm{f}_n^2\Bigr] 
-	\end{equation}
-	with probability at least $1 - 5 \delta$.
+	\norm{f}_n^2 \geq \delta \cdot \Ebb\Bigl[\norm{f}_n^2\Bigr].
+	\end{equation}
 \end{lemma}
@@
 then \eqref{eqn:empirical_norm_sobolev} follows from the Paley-Zygmund inequality (Lemma~\ref{lem:paley_zygmund}).
-Since $p \leq p_{\max}$ is uniformly bounded, we can relate $\mathbb{E}\bigl[\norm{f}_n^4\bigr]$ to the $\Leb^4$ norm,
-\begin{equation*}
-\mathbb{E}\bigl[\norm{f}_n^4\bigr] = \frac{(n-1)}{n}\left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + \frac{\mathbb{E}\Bigl[\bigl(f(x_1)\bigr)^4\Bigr]}{n} \leq \left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + p_{\max}^2\frac{\norm{f}_{\Leb^4}^4}{n}.
-\end{equation*}
-We will use a Sobolev inequality to relate $\norm{f}_{\Leb^4(\Xset)}$ to $\norm{f}_{H^1(\Xset)}$.
-The nature of this inequality depends on the relationship between $s$ and $d$ (see Theorem 6 in Section 5.6.3 of \cite{evans10} for a formal statement), so from this point on we divide our analysis into three cases: (i) the case where $2 > d$, (ii) the case where $2 < d$, and (iii) the borderline case $2 = d$.
-
-\textit{Case 1: $2s > d$.}
-When $2 > d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in C^{\gamma}(\overline{\Xset})$ for some $\gamma > 0$ which depends on $d$, with the accompanying estimate
-\begin{equation*}
-\sup_{x \in \Xset} \abs{f(x)} \leq \norm{f}_{C^{\gamma}(\Xset)} \leq c \norm{f}_{H^1(\Xset)}.
-\end{equation*}
-Therefore,
-\begin{align*}
-\norm{f}_{\Leb^4(\Xset)}^4 & = \int_{\Xset} [f(x)]^4 \,dx \\
-& \leq \left(\sup_{x \in \Xset} \abs{f(x)}\right)^2 \cdot \int_{\Xset} [f(x)]^2 \,dx \\
-& \leq c \norm{f}_{H^1(\Xset)}^2 \cdot \norm{f}_{\Leb^2(\Xset)}^2.
+Since $p \leq p_{\max}$ is uniformly bounded, we can relate $\mathbb{E}\bigl[\norm{f}_n^4\bigr]$ to the $\Leb^4(\Xset)$-norm,
+\begin{equation*}
+\mathbb{E}\bigl[\norm{f}_n^4\bigr] = \frac{(n-1)}{n}\left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + \frac{\mathbb{E}\Bigl[\bigl(f(x_1)\bigr)^4\Bigr]}{n} \leq \left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + p_{\max}\frac{\norm{f}_{\Leb^4(\Xset)}^4}{n}.
+\end{equation*}
+We will use the Sobolev inequalities as a tool to show that $\|f\|_{\Leb^4(\Xset)}/n \leq \bigl(\mathbb{E}[\|f\|_n^2]\bigr)^2/(b^2p_{\max})$, whence the claim of the Lemma is shown.
+The nature of the inequalities we use depend on the value of $d$.
+In particular, we will use the following relationships between norms:
+\begin{equation*}
+\begin{rcases*}
+\sup_{x \in \mc{X}}|f(x)|,& \textrm{$d = 1$}\\
+\|f\|_{\Leb^{q}(\mc{X})},& \textrm{$d = 2$, for all $0 < q < \infty$} \\
+\|f\|_{\Leb^{q}(\mc{X})},& \textrm{$d \geq 3$, for all $0 < q \leq 2d/(d - 2)$}
+\end{rcases*}
+\leq C_7 \cdot M.
+\end{equation*}
+(See Theorem 6 in Section 5.6.3 of \citet{evans10} for a complete statement and proof of the various Sobolev inequalities.)
+
+As a result, we divide our analysis into three cases: (i) the case where $d < 2$, (ii) the case where $d > 2$, and (iii) the borderline case $d = 2$.
+
+\textit{Case 1: $d < 2$.}
+The $\Leb^4(\Xset)$-norm of $f$ can be bounded in terms of the $\Leb^2(\Xset)$ norm,
+\begin{align*}
+\norm{f}_{\Leb^4(\Xset)}^4 & \leq \left(\sup_{x \in \Xset} \abs{f(x)}\right)^2 \cdot \int_{\Xset} [f(x)]^2 \,dx \leq C_7^2 M^2 \cdot \|f\|_{\Leb^2(X)}^2.
 \end{align*}
@@
 \begin{equation*}
-\norm{f}_{\Leb^2(\Xset)}^2 \geq c_1^2 \cdot b^2 \cdot \norm{f}_{H^{1}(\Xset)}^2 \cdot \frac{1}{n},
+\norm{f}_{\Leb^2(\Xset)}^2 \geq C_6^2 \cdot b^2 \cdot M^2 \cdot \frac{1}{n},
 \end{equation*}
@@
 \begin{equation*}
-p_{\max}^2\frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c \norm{f}_{H^1(\Xset)}^2 \cdot \frac{\norm{f}_{\Leb^2(\Xset)}^4}{n \norm{f}_{\Leb^2(\Xset)}^2} \leq c \frac{\norm{f}_{\Leb^2(\Xset)}^4}{c_1^2 b^2} \leq \frac{\Ebb\bigl[\norm{f}_n^2\bigr]}{b^2},
-\end{equation*}
-where the last inequality follows by taking $c_1$ sufficiently large.
-
-\textit{Case 2: $2 < d$.}
-When $2 < d$, since $\Xset$ is a Lipschitz domain the Sobolev inequality establishes that $f \in \Leb^q(\Xset)$ for $q = 2d/(d - 2)$, and moreover that
-\begin{equation*}
-\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{H^1(\Xset)}.
-\end{equation*}
-Since $4 = 2\theta + (1 - \theta)q$ for $\theta = 2 - d/2$, Lyapunov's inequality implies
-\begin{equation*}
-\norm{f}_{\Leb^4(\Xset)}^4 \leq \norm{f}_{\Leb^2}^{2\theta} \cdot \norm{f}_{\Leb^q(\Xset)}^{(1 - \theta)q} \leq c \norm{f}_{\Leb^2(\Xset)}^{4} \cdot \left(\frac{\norm{f}_{H^1(\Xset)}}{\norm{f}_{\Leb^2(\Xset)}}\right)^{d}.
-\end{equation*}
-By assumption, $\norm{f}_{\Leb^2(\Xset)} \geq c_1 b \norm{f}_{H^1(\Xset)} n^{-1/d}$, and therefore
-\begin{equation*}
-p_{\max}^2 \frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq c\norm{f}_{\Leb^2(\Xset)}^4 \left(\frac{\norm{f}_{H^1(\Xset)}}{n^{1/d}\norm{f}_{\Leb^2(\Xset)}}\right)^{d} \leq \frac{c\norm{f}_{\Leb^2(\Xset)}^4}{c_1b^{d}} \leq \frac{\norm{f}_{\Leb^2(\Xset)}^4}{b^2}.
-\end{equation*}
-where the last inequality follows when $c_1$ is sufficiently large, and keeping in mind that $d > 2$ and $b \geq 1$.
-
-\textit{Case 3: $2 = d$.}
-Assume $f$ satisfies~\eqref{eqn:paley_zygmund_1} for a given $0 < a < 1$.
-When $2 = d$, since $\Xset$ is a Lipschitz domain we have that $f \in L^q(\Xset)$ for any $q < \infty$, with the accompanying estimate
-\begin{equation*}
-\norm{f}_{\Leb^q(\Xset)} \leq c \norm{f}_{H^1(\Xset)}.
-\end{equation*}
-In particular the above holds for $q = 2/(1 - a)$ when $1/2 < a < 1$, and for any $q > 4$ when $0 < a < 1/2$. Using Lyapunov's inequality as in the previous case then implies the desired result.
+p_{\max} \frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq C_7^2 M^2p_{\max} \cdot \frac{\norm{f}_{\Leb^2(\Xset)}^2}{n} \leq \frac{C_7p_{\max}}{C_6^2 b^2} \norm{f}_{\Leb^2(\Xset)}^4 \leq \frac{\Ebb\bigl[\norm{f}_n^2\bigr]}{b^2},
+\end{equation*}
+where the last inequality follows by taking $C_6 \geq C_7 \sqrt{p_{\max}/p_{\min}}$.
+
+\textit{Case 2: $d > 2$.}
+Let $\theta = 2 - d/2$ and $q = 2d/(d - 2)$. Noting that $4 = 2\theta + (1 - \theta)q$, Lyapunov's inequality implies
+\begin{equation*}
+\norm{f}_{\Leb^4(\Xset)}^4 \leq \norm{f}_{\Leb^2(\Xset)}^{2\theta} \cdot \norm{f}_{\Leb^q(\Xset)}^{(1 - \theta)q} \leq \norm{f}_{\Leb^2(\Xset)}^{4} \cdot \left(\frac{C_7\norm{f}_{H^1(\Xset)}}{\norm{f}_{\Leb^2(\Xset)}}\right)^{d}.
+\end{equation*}
+By assumption, $\norm{f}_{\Leb^2(\Xset)} \geq C_6 b \norm{f}_{H^1(\Xset)} n^{-1/d}$, and therefore
+\begin{equation*}
+p_{\max} \frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq \norm{f}_{\Leb^2(\Xset)}^4p_{\max} \cdot \left(\frac{C_7\norm{f}_{H^1(\Xset)}}{n^{1/d}\norm{f}_{\Leb^2(\Xset)}}\right)^{d} \leq \frac{C_7^dp_{\max}\norm{f}_{\Leb^2(\Xset)}^4}{C_6^db^{d}} \leq \frac{\Ebb\bigl[\norm{f}_n^2\bigr]}{b^2}.
+\end{equation*}
+where the last inequality follows by taking $C_6 \geq C_7(p_{\max}/p_{\min})^{1/d}$, and keeping in mind that $d > 2$ and $b \geq 1$.
+
+\textit{Case 3: $d = 2$.}
+Fix $t \in (1/2,1)$, and suppose that
+\begin{equation}
+\label{pf:empirical_norm_sobolev_1}
+\|f\|_{\Leb^2(\Xset)} \geq \frac{C_6 M}{\delta} \cdot n^{-t/2}.
+\end{equation} 
+Putting $q = 2/(1 - t)$, we have that $\|f\|_{\Leb^{q}(\mc{X})} \leq C_7 \cdot M$, and it follows from derivations similar to those in Case 2 that $\|f\|_{\Leb^4(\Xset)}/n \leq \bigl(\mathbb{E}[\|f\|_n^2]\bigr)^2/(b^2p_{\max})$ when $C_6 \geq C_7 \sqrt{p_{\max}/p_{\min}}$.
+
+Now, suppose $f \in \Leb^4(\Xset)$ satisfies~\eqref{pf:empirical_norm_sobolev_1} only when $t = 1$.
+For each $k = 1,2,\ldots$ let $f_k := n^{1/(2k)}f$, so that each $f_k$ satisfies~\eqref{pf:empirical_norm_sobolev_1} with respect to $t = 1 - 1/k$.
+Clearly $\|f_k - f\|_{\Leb^4(\Xset)} \to 0$ as $k \to \infty$, and therefore
+\begin{equation*}
+\frac{1}{n}\|f\|_{\Leb^4(\Xset)} = \frac{1}{n}\lim_{k \to \infty} \|f_k\|_{\Leb^4(\Xset)} \leq \frac{1}{b^2p_{\max}} \lim_{k \to \infty} \bigl(\mathbb{E}[\|f_k\|_n^2]\bigr)^2 =  \frac{1}{b^2p_{\max}}\bigl(\mathbb{E}[\|f\|_n^2]\bigr)^2.
+\end{equation*}
+This establishes the claim when $d = 2$, and completes the proof of Lemma~\ref{lem:empirical_norm_sobolev}.
 
@@
 \label{sec:manifold}
-In this section, we restate a few results of \citep{trillos2019,calder2019}, which are analogous to Lemmas~\ref{lem:graph_sobolev_seminorm} and~\ref{lem:neighborhood_eigenvalue} but cover the case where $\Xset$ is an $m$-dimensional submanifold without boundary.
+In this section, we restate a few results of \citet{trillos2019,calder2019}, which are analogous to Lemmas~\ref{lem:graph_sobolev_seminorm} and~\ref{lem:neighborhood_eigenvalue} but cover the case where $\Xset$ is an $m$-dimensional submanifold without boundary.
 As such, the results in this section will hold under the assumption~\ref{asmp:domain_manifold}.
-We refer to~\citep{trillos2019,calder2019} for the proofs of these results.
-
-Proposition~\ref{prop:garciatrillos19_1} follows from Lemma~5 of~\citep{trillos2019} and Markov's inequality.
-\begin{proposition}[]
+We refer to~\citet{trillos2019,calder2019} for the proofs of these results.
+
+Proposition~\ref{prop:garciatrillos19_1} follows from Lemma~5 of~\citet{trillos2019} and Markov's inequality.
+\begin{proposition}
 	\label{prop:garciatrillos19_1}
@@
 	\begin{equation*}
-	f^T \Lap_{n,r} f \leq \frac{C}{\delta} n^2 r^{m + 2} |f|_{H^1(\Xset)}^2
+	f^{\top} \Lap_{n,r} f \leq \frac{C}{\delta} n^2 r^{m + 2} |f|_{H^1(\Xset)}^2
 	\end{equation*}
@@
 	\begin{equation*}
-	n^{-1/m} < \wt{\delta} \leq \frac{1}{4}r~~\textrm{and}~~C(\theta + \wt{\delta}) \leq \frac{1}{2}p_{\min}~~\textrm{and}~~r\leq\min\{c_3,1\}.
+	n^{-1/m} < \wt{\delta} \leq \frac{1}{4}r~~\textrm{and}~~C(\theta + \wt{\delta}) \leq \frac{1}{2}p_{\min}~~\textrm{and}~~C_4\bigl(\log(n)/n\bigr)^{1/m}\leq r \leq\min\{c_4,1\}.
 	\end{equation*}
@@
 
-\begin{proposition}[c.f Theorem 2.4 of~\citep{calder2019}]
+\begin{proposition}[\textbf{c.f Theorem 2.4 of~\citet{calder2019}}]
 	\label{prop:calder19_1}
@@
 
-Proposition~\ref{prop:calder19_2} follows from Lemma 3.1 of \citep{calder2019}, along with a union bound.
+Proposition~\ref{prop:calder19_2} follows from Lemma 3.1 of \citet{calder2019}, along with a union bound.
 \begin{proposition}
@@
 We are now in a position to prove Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing_manifold}, as well as a few other claims from our main text.
-We divide our proofs into estimation results and testing results.
-Throughout, it will be convenient for us to deal with the normalization $\wt{\rho} = \rho nr^{d + 2}$.
+In Section~\ref{subsec:laplacian_smoothing_estimation1_pf} we prove all of our results regarding estimation and in Section~\ref{subsec:laplacian_smoothing_testing_pf} we prove all of our results regarding testing; in Section~\ref{subsec:convenient_estimate}, Lemmas~\ref{lem:variance_term_estimation} and~\ref{lem:variance_term_testing}, we provide some useful estimates on a particular pair of sums that appear repeatedly in our proofs.
+Throughout, it will be convenient for us to deal with the normalization $\wt{\rho} := \rho nr^{d + 2}$.
+We note that in each of our Theorems, the prescribed choice of $\rho$ will always result in $\wt{\rho} \leq 1$.
 
@@
 \paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}.}
-
-From Lemmas~\ref{lem:graph_sobolev_seminorm},~\ref{lem:neighborhood_eigenvalue}, and~\ref{lem:ls_fixed_graph_estimation}, we have that with probability at least $1 - \delta - C_1n\exp(-c_1nr^d) - \exp(-\sum_{k = 1}^{n}(\rho \lambda_k + 1)^{-2})$,
+We have shown that the inequalities~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} are satisfied with probability at least $1 - \delta - C_1n\exp(-c_1nr^d)$.
+From~\eqref{eqn:neighborhood_eigenvalue} and Lemma~\ref{lem:variance_term_estimation}, we have that
+\begin{equation*}
+\sum_{k = 1}^{n}\biggl(\frac{1}{\rho \lambda_k + 1}\biggr)^{2} \geq 1 + \sum_{k = 2}^{n}\biggl(\frac{1}{C_3 \wt{\rho} k^{2/d} + 1}\biggr)^{2} \geq 1 + \frac{1}{4} \min\Bigl\{\floor{\frac{\wt{\rho}^{-d/2}}{C_3^{d/2}}} - 1 \vee 0,n\Bigr\} \geq \min\Bigl\{\frac{\wt{\rho}^{-2/d}}{8C_3^{2/d}}, \frac{n}{4} \Bigr\}
+\end{equation*}
+Therefore by Lemma~\ref{lem:ls_fixed_graph_estimation} along with~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue}, with probability at least $1 - \delta - C_1n\exp(-c_1nr^d) - \exp(-\min\{\wt{\rho}^{-d/2}/C_3^{d/2},n\}/8)$ it follows that,
 \begin{align}
-\|\wh{f} - f_0\|_n^2 & \leq \frac{C_3}{\delta} \wt{\rho} M^2 + \frac{10}{n} \sum_{k = 1}^{n} \Biggl(\frac{1}{\wt{\rho}\min\{k^{2/d},r^{-2}\} + 1}\Biggr)^2 \nonumber \\
-& \leq \frac{C_3}{\delta} \wt{\rho} M^2 + \frac{10}{n} \sum_{k = 1}^{n} \biggl(\frac{1}{\wt{\rho}k^{2/d} + 1}\biggr)^2 + \frac{10r^4}{\wt{\rho}^2} \label{pf:laplacian_smoothing_estimation1_1}
+\|\wh{f} - f_0\|_n^2 & \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{n}\sum_{k = 2}^{n} \Biggl(\frac{1}{c_3 \wt{\rho}\min\{k^{2/d},r^{-2}\} + 1}\Biggr)^2 \nonumber \\
+& \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{n}\sum_{k = 1}^{n} \biggl(\frac{1}{c_3 \wt{\rho}k^{2/d} + 1}\biggr)^2 + \frac{10r^4}{c_3^2 \wt{\rho}^2}.
+\label{pf:laplacian_smoothing_estimation1_1}
 \end{align}
-where $C_3 := (1 + L_pc_0)^2 \sigma_K$.
-Setting $\wt{\rho} = M^{-4/(2 + d)}n^{-2/(2 + d)}$ as prescribed in Theorem~\ref{thm:laplacian_smoothing_estimation1}, from Lemma~\ref{lem:variance_term_estimation}:
-\begin{align*}
-\|\wh{f} - f_0\|_n^2 & \leq\frac{C_3}{\delta} \wt{\rho} M^2 + \frac{20\wt{\rho}^{-d/2}}{n}  + \frac{10r^4}{\wt{\rho}^2} \\
-& = \biggl(\frac{C_3}{\delta} + 10\biggr)M^{2d/(2+d)} n^{-2/(2 + d)} + 10r^4M^{8/(2 + d)}n^{4/(2 + d)} \\
-& \leq \biggl(\frac{C_3}{\delta} + 20\biggr)M^{2d/(2+d)} n^{-2/(2 + d)}
-\end{align*}
-with the last inequality following from~\ref{asmp:ls_kernel_radius_estimation}. Finally, observe that by Lemmas~\ref{lem:neighborhood_eigenvalue} and~\ref{lem:variance_term_estimation},
-\begin{equation*}
-\exp\biggl(-\sum_{k = 1}^{n}(\rho \lambda_k + 1)^{-2}\biggr) \leq \exp\biggl(-\sum_{k = 1}^{n}(\wt{\rho} k^{2/d} + 1)^{-2}\biggr) \leq \exp\biggl(-\frac{1}{4}\min\Bigl\{\floor{\wt{\rho}^{-d/2}},n\Bigr\}\biggr),
-\end{equation*}
-completing the proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}.
+Setting $\wt{\rho} = M^{-4/(2 + d)}n^{-2/(2 + d)}$ as prescribed in Theorem~\ref{thm:laplacian_smoothing_estimation1},  we have from Lemma~\ref{lem:variance_term_estimation} that
+\begin{align*}
+\|\wh{f} - f_0\|_n^2 & \leq\frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{20\wt{\rho}^{-d/2}}{c_3^{d/2}n}  + \frac{10r^4}{c_3^2 \wt{\rho}^2} \\
+& = \biggl(\frac{C_2}{\delta} + \frac{20}{c_3^{d/2}}\biggr)M^{2d/(2+d)} n^{-2/(2 + d)} + \frac{10}{n} + \frac{10}{c_3^2}r^4M^{8/(2 + d)}n^{4/(2 + d)} \\
+& \leq \biggl(\frac{C_2}{\delta} + \frac{20}{c_3^{d/2}} + 10 + \frac{10}{c_3^2}\biggr)M^{2d/(2+d)} n^{-2/(2 + d)}
+\end{align*}
+with the last inequality following from~\ref{asmp:ls_kernel_radius_estimation} and the assumption $M \geq n^{-1/2}$.
+This completes the proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}.
+When $M < n^{-1/2}$, one can set $\wt{\rho} = 1$ and trace through the previous calculations to determine that $\|\wh{f} - f_0\|_n^2 \leq C/n$ for an appropriate choice of $C$.
 
@@
 \begin{equation*}
-\|\wh{f} - f_0\|_n^2 \leq \frac{C_3}{\delta} \wt{\rho} M^2 + \frac{10}{n\wt{\rho}^2} +  \frac{10 \log n}{n\wt{\rho}^2} + \frac{10r^4}{\wt{\rho}^2}
-\end{equation*}
-Setting $r = (C_4\log(n)/n)^{1/4}$, we obtain
-\begin{equation*}
-\|\wh{f} - f_0\|_n^2 \leq \frac{C_3}{\delta} \wt{\rho} M^2 + \frac{10}{n\wt{\rho}^2} +  \frac{10 \log n}{n\wt{\rho}^2} + \frac{10C_4\log n}{n\wt{\rho}^2},
+\|\wh{f} - f_0\|_n^2 \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} +  \frac{10}{nc_3^2\wt{\rho}^2} +  \frac{10 \log n}{nc_3^2\wt{\rho}^2} + \frac{10r^4}{c_3^2\wt{\rho}^2}
+\end{equation*}
+Setting $r = (C_0\log(n)/n)^{1/4}$, we obtain
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} +  \frac{10}{nc_3^2\wt{\rho}^2} +  \frac{10 \log n}{nc_3^2\wt{\rho}^2} + \frac{10C_0\log n}{nc_3^2\wt{\rho}^2},
 \end{equation*}
@@
 \begin{equation*}
-\|\wh{f} - f_0\|_n^2 \leq \biggl(\frac{C_3}{\delta} + 20 + 10C_4\biggr) M^{4/3} \biggl(\frac{\log n}{n}\biggr)^{1/3}.
+\|\wh{f} - f_0\|_n^2 \leq \biggl(\frac{C_2}{\delta} + \frac{20}{c_3^2} + \frac{10C_0}{c_3^2}\biggr) M^{4/3} \biggl(\frac{\log n}{n}\biggr)^{1/3} + \frac{10}{n}.
 \end{equation*}
@@
 \begin{equation*}
-\|\wh{f} - f_0\|_n^2 \leq \frac{C_3}{\delta} \wt{\rho} M^2 + \frac{10}{n\wt{\rho}^{d/2}} +  \frac{10}{n^{4/d}\wt{\rho}^2} + \frac{10r^4}{\wt{\rho}^2}
-\end{equation*}
-Setting $r = (C_4\log n/n)^{1/d}$, we obtain
-\begin{equation*}
-\|\wh{f} - f_0\|_n^2 \leq \frac{C_3}{\delta} \wt{\rho} M^2 + \frac{10}{n\wt{\rho}^{d/2}} +  \frac{10}{n^{4/d}\wt{\rho}^2} + \frac{10C_4(\log n)^{4/d}}{n^{4/d}\wt{\rho}^2},
+\|\wh{f} - f_0\|_n^2 \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{nc_3^{d/2}\wt{\rho}^{d/2}} +  \frac{10}{n^{4/d}\wt{\rho}^2c_3^2} + \frac{10r^4}{\wt{\rho}^2c_3^2}
+\end{equation*}
+Setting $r = (C_0\log n/n)^{1/d}$, we obtain
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{n\wt{\rho}^{d/2}c_3^{d/2}} +  \frac{10}{n^{4/d}\wt{\rho}^2c_3^2} + \frac{10C_0^{4/d}(\log n)^{4/d}}{n^{4/d}\wt{\rho}^2c_3^2},
 \end{equation*}
@@
 \begin{equation*}
-\|\wh{f} - f_0\|_n^2 \biggl(\frac{C_3}{\delta} + 10 + 10C_4\biggr)M^{4/d}\biggl(\frac{\log(n)}{n^{1/3}}\biggr)^{4/d} + 10M^{d/3}n^{-1/3}.
+\|\wh{f} - f_0\|_n^2 \leq \biggl(\frac{C_2}{\delta} + \frac{10}{c_3^2} + \frac{10C_0^{4/d}}{c_3^2}\biggr)M^{4/d}\biggl(\frac{\log(n)}{n^{1/3}}\biggr)^{4/d} + \frac{10}{c_3^{d/2}}M^{d/3}n^{-1/3} + \frac{10}{n}.
 \end{equation*}
@@
 \paragraph{Bounds on $\Leb^2(\Xset)$ error under Lipschitz assumption.}
-Suppose $f_0$ has bounded Lipschitz constant.
-Let $V_1,\ldots,V_n$ denote the Voronoi tesselation of $\Xset$ with respect to $X_1,\ldots,X_n$.
-Extend $\wh{f}$ over $\Xset$ by taking it piecewise constant over the Voronoi cells, i.e.
+Let $V_1,\ldots,V_n$ denote the Voronoi tesselation of $\Xset$ with respect to $X_1,\ldots,X_n$. Extend $\wh{f}$ over $\Xset$ by taking it piecewise constant over the Voronoi cells, i.e.
 \begin{equation*}
@@
 Note that $n^{-2/d} \ll n^{-2/(2 +d)}$.
-Therefore Proposition~\ref{prop:out_of_sample_error} together with Theorem~\ref{thm:laplacian_smoothing_estimation1} implies that with high probability, $\wh{f}$ achieves the nearly-optimal (up to a factor of $\log n$) estimation out-of-sample error---that is, $\|\wh{f} - f_0\|_{\Leb^2(\Xset)}^2 \leq C \log(n) M^{2d/(2 + d)}n^{-2/(2+d)}$---as long as $M \leq Cn^{1/d}$.
+Therefore Proposition~\ref{prop:out_of_sample_error} together with Theorem~\ref{thm:laplacian_smoothing_estimation1} implies that with high probability, $\wh{f}$ achieves the nearly-optimal (up to a factor of $\log n$) estimation rates out-of-sample error---that is, $\|\wh{f} - f_0\|_{\Leb^2(\Xset)}^2 \leq C \log(n) M^{2d/(2 + d)}n^{-2/(2+d)}$---as long as $M \leq Cn^{1/d}$.
 \paragraph{Proof of Proposition~\ref{prop:out_of_sample_error}.}
@@
 \label{pf:out_of_sample_error_0}
-\|\wh{f} - f\|_{\Leb^2(\Xset)}^2 \leq 2 \sum_{i = 1}^{n} \vol(V_i) \Bigl(\wh{f}(X_i) - f_0(X_i)\Bigr)^2 + 2 \sum_{i = 1}^{n} \Bigl(\mathrm{diam}(V_i)\Bigr)^2
+\|\wh{f} - f\|_{\Leb^2(\Xset)}^2 \leq 2 \sum_{i = 1}^{n} \vol(V_i) \Bigl(\wh{f}(X_i) - f_0(X_i)\Bigr)^2 + 2 M^2 \sum_{i = 1}^{n} \Bigl(\mathrm{diam}(V_i)\Bigr)^2
 \end{equation}
@@
 
-Now we will use some results of \citep{chaudhuri2010} regarding uniform concentration of empirical counts, to upper bound $\diam(V_i)$ Set
-\begin{equation*}
-\varepsilon_n := \biggl(\frac{2C_{o}\log(1/\delta)d\log n}{\nu_dp_{\min}a_3n}\biggr)^{1/d}
-\end{equation*}
-and note that for $n$ sufficiently large, $\varepsilon_n \leq c_0$.
-Thus by~\eqref{eqn:integral_boundary} we have that for every $x \in \Xset$, $P(B(x,\varepsilon_n)) \geq 2C_{o}\log(1/\delta)d\frac{\log n}{n}$.
-Consequently, by Lemma~16 of \citep{chaudhuri2010} it holds that with probability at least $1 - \delta$,
+Now we will use some results of \citet{chaudhuri2010} regarding uniform concentration of empirical counts, to upper bound $\diam(V_i)$ Set
+\begin{equation*}
+\varepsilon_n := \biggl(\frac{2C_{o}\log(1/\delta)d\log n}{\nu_dp_{\min}a_3n}\biggr)^{1/d},
+\end{equation*}
+where $C_{o}$ is a constant given in Lemma~16 of \citet{chaudhuri2010}.
+Note that for $n$ sufficiently large, $\varepsilon_n \leq c_0$, and therefore by~\eqref{eqn:integral_boundary} we have that for every $x \in \Xset$, $P(B(x,\varepsilon_n)) \geq 2C_{o}\log(1/\delta)d\frac{\log n}{n}$.
+Consequently, by Lemma~16 of \citet{chaudhuri2010} it holds that with probability at least $1 - \delta$,
 \begin{equation}
@@
 \end{equation}
-But if~\eqref{pf:out_of_sample_error_1} is true, it must also hold that for each $i = 1,\ldots,n$ and for every $x \in V_i$, the distance $\|x - X_i\| \leq \varepsilon_n$, and so
-\begin{equation*}
-\max_{i = 1,\ldots,n} \mathrm{diam}(V_i) \leq 2\varepsilon_n.
-\end{equation*}
+But if~\eqref{pf:out_of_sample_error_1} is true, it must also be true that for each $i = 1,\ldots,n$ and for every $x \in V_i$, the distance $\|x - X_i\| \leq \varepsilon_n$.
+Thus by the triangle inequality, $\max_{i = 1,\ldots,n} \mathrm{diam}(V_i) \leq 2\varepsilon_n$.
 Plugging back in to~\eqref{pf:out_of_sample_error_0}, and using the upper bound volume $\vol(V_i) \leq \nu_d \bigl(\mathrm{diam}(V_i)\bigr)^d$, we obtain the desired upper bound on $\|\wh{f} - f\|_{\Leb^2(\Xset)}^2$.
@@
 \paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_estimation_manifold}.}
-The proof of Theorem~\ref{thm:laplacian_smoothing_estimation_manifold} follows exactly the same steps as the proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}, replacing the references to Lemma~\ref{lem:graph_sobolev_seminorm} and~\ref{lem:neighborhood_eigenvalue} by references to Propositions~\ref{prop:garciatrillos19_1} and~\eqref{eqn:neighborhood_eigenvalue_manifold}, and the ambient dimension $d$ by the intrinsic dimension $m$.
+The proof of Theorem~\ref{thm:laplacian_smoothing_estimation_manifold} follows exactly the same steps as the proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}, replacing the references to Lemma~\ref{lem:graph_sobolev_seminorm} and~\ref{lem:neighborhood_eigenvalue} by references to Proposition~\ref{prop:garciatrillos19_1} and~\eqref{eqn:neighborhood_eigenvalue_manifold}, and the ambient dimension $d$ by the intrinsic dimension $m$.
 
 \subsection{Proofs of testing results}
+\label{subsec:laplacian_smoothing_testing_pf}
 
 \paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_testing}.}
-We would like to invoke Lemma~\ref{lem:ls_fixed_graph_testing}, and in order to do so, we must show that~\eqref{eqn:ls_fixed_graph_testing_critical_radius} is satisfied with respect to $G = G_{n,r}$.
-On the one hand, by Lemmas~\ref{lem:graph_sobolev_seminorm} and~\ref{lem:neighborhood_eigenvalue}, with probability at least $1 - b - C_1n\exp(-c_1nr^d)$,
-\begin{align*}
-\frac{2 \rho}{n} \bigl(f_0^T \Lap f_0\bigr) + \frac{4b}{n} \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{1/2} & \leq C_3b\wt{\rho}M^2 + \frac{4b}{n}\Biggl[\biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{1/2} + \frac{r^8}{\wt{\rho}^4n^{1/2}}\Biggr] \\
-& \leq C_3b\wt{\rho}M^2 + \frac{4b}{n}\biggl(\sqrt{2}\wt{\rho}^{-d/4} + \frac{r^8}{\wt{\rho}^4n^{1/2}}\biggr) \\ 
-& \leq \Bigl(C_3 + 12\Bigr)bM^{2d/(4 + d)} n^{-4/(4 + d)}
-\end{align*}
-where the second inequality comes from Lemma~\ref{lem:variance_term_testing}, and in the third inequality we set $\wt{\rho} = M^{-8/(4 + d)}n^{-4/(4 + d)}$ as prescribed by Theorem~\ref{thm:laplacian_smoothing_testing}.
-
-On the other hand, by Lemma~\ref{lem:empirical_norm_sobolev}, with probability at least $1 - 5/b$,
-\begin{equation*}
-\|f\|_n^2 \geq \frac{p_{\min}}{b} \|f\|_{\Leb^2(\Xset)}^2 \geq CbM^{2d/(4 + d)}n^{-4/(4 + d)}
-\end{equation*}
-and so choosing $C$ appropriately in~\eqref{eqn:laplacian_smoothing_testing}, ~\eqref{eqn:ls_fixed_graph_testing_critical_radius} is verified.
-
-By these derivations along with Lemma~\ref{lem:ls_fixed_graph_testing}, the Laplacian smoothing test $\wh{\varphi}$ has Type II error upper bounded by
+Recall that we have shown that the inequalities~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} are satisfied with probability at least $1 - \delta - C_1n\exp(-c_1nr^d)$.
+
+Now, we would like to invoke Lemma~\ref{lem:ls_fixed_graph_testing}, and in order to do so, we must show that the inequality~\eqref{eqn:ls_fixed_graph_testing_critical_radius} is satisfied with respect to $G = G_{n,r}$.
+First, we upper bound the right hand side of this inequality.
+By~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue},
+\begin{align*}
+\frac{2 \rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{4b}{n} \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{1/2} & \leq C_2b\wt{\rho}M^2 + \frac{4b}{n}\Biggl[1 + \biggl(\sum_{k = 2}^{n} \frac{1}{(c_3\wt{\rho}k^{2/d} + 1)^4} \biggr)^{1/2} + \frac{r^4n^{1/2}}{c_3^2\wt{\rho}^2}\Biggr] \\
+& \leq C_2b\wt{\rho}M^2 + \frac{4b}{n}\biggl(1 + \frac{\sqrt{2}}{c_3^{d/4}}\wt{\rho}^{-d/4} + \frac{r^4n^{1/2}}{c_3^2\wt{\rho}^2}\biggr) \\ 
+& \leq \Bigl(C_2 + 4 + \frac{4\sqrt{2}}{c_3^{d/4}} + \frac{4}{c_3^2}\Bigr)bM^{2d/(4 + d)} n^{-4/(4 + d)}
+\end{align*}
+where the second inequality is justfied by Lemma~\ref{lem:variance_term_testing}, and in the third inequality we set $\wt{\rho} = M^{-8/(4 + d)}n^{-4/(4 + d)}$ as prescribed by Theorem~\ref{thm:laplacian_smoothing_testing}, and use both the upper bound on $r$ assumed in~\ref{asmp:ls_kernel_radius_testing} as well as the fact that $M \geq n^{-1/2}$.
+
+Next we lower bound the left hand side of the inequality~\eqref{eqn:ls_fixed_graph_testing_critical_radius}---i.e.
+we lower bound the empirical norm $\|f_0\|_n^2$---using Lemma~\ref{lem:empirical_norm_sobolev}.
+Recall that by assumption, $M \leq M_{\max}(d)$.
+Therefore, taking $C \geq C_6$ in~\eqref{eqn:laplacian_smoothing_testing}, we have that~\eqref{eqn:empirical_norm_sobolev_1} is satisfied, so that it follows from~\eqref{eqn:empirical_norm_sobolev} that
+\begin{equation*}
+\|f\|_n^2 \geq \frac{\|f\|_{P}^2}{b} \geq \frac{p_{\min}}{b} \|f\|_{\Leb^2(\Xset)}^2 \geq CbM^{2d/(4 + d)}n^{-4/(4 + d)},
+\end{equation*}
+with probability at least $1 - 5/b$.
+Taking $C \geq C_2 + 4 + (4\sqrt{2})/c_3^{d/4} + 4/c_3^2$ in~\eqref{eqn:laplacian_smoothing_testing} thus implies~\eqref{eqn:ls_fixed_graph_testing_critical_radius}, and we may therefore use Lemma~\ref{lem:ls_fixed_graph_testing} to upper bound the type II error the Laplacian smoothing test $\wh{\varphi}$.
+Observe that by~\eqref{eqn:neighborhood_eigenvalue} and the lower bound in Lemma~\ref{lem:variance_term_testing},
+\begin{equation*}
+\sum_{k = 1}^{n} \biggl(\frac{1}{\rho \lambda_k + 1}\biggr)^4 \geq 1 + \sum_{k = 2}^{n}\biggl(\frac{1}{C_3\wt{\rho}k^{2/d} + 1}\biggr)^4 \geq 1 + \frac{1}{16}\min\biggl\{\floor{\frac{\wt{\rho}^{-d/2}}{C_3^{d/2}}} - 1 \vee 0, n \biggr\} \geq \min\biggl\{\frac{\wt{\rho}^{-d/2}}{32 C_3^{d/2}}, \frac{n}{16} \biggr\}.
+\end{equation*}
+We conclude that
 \begin{align*}
 \Pbb_{f_0}\bigl(\wh{T} \leq \wh{t}_b\bigr) & \leq \frac{6}{b} + \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2} + C_1n\exp(-c_1nr^d)\\
-& \leq \frac{6}{b} + \frac{4}{b^2} + \frac{32}{b} \max\Bigl\{\floor{\wt{\rho}^{-d/4}},n^{-1/2}\Bigr\} + C_1n\exp(-c_1nr^d)
+& \leq \frac{10}{b} + \frac{32\sqrt{2}}{b} \max\Bigl\{ C_3^{d/4}\wt{\rho}^{d/4} ,n^{-1/2}\Bigr\} + C_1n\exp(-c_1nr^d),
 \end{align*}
@@
 \paragraph{Proof of~\eqref{eqn:laplacian_smoothing_testing_low_smoothness}.}
-When $\rho = 0$, the Laplacian smoother $\wh{f} = Y$, the test statistic $\wh{T} = \frac{1}{n}\|Y\|_2^2$, and the threshold $\wh{t}_b = 1 + 2bn^{-1/2}$. The expectation of $\wh{T}$ is 
+When $\rho = 0$, the Laplacian smoother $\wh{f} = \mathbf{Y}$, the test statistic $\wh{T} = \frac{1}{n}\|\mathbf{Y}\|_2^2$, and the threshold $\wh{t}_b = 1 + 2bn^{-1/2}$.
+The expectation of $\wh{T}$ is
 \begin{equation*}
@@
 \end{equation*}
-so that $E[\wh{T}] - \wh{t}_b \geq E[\wh{T}]/2$. Hence, by Chebyshev's inequality
-\begin{align*}
-\mathbb{P}_{f_0}\Bigl(\wh{T} \leq \wh{t}_b \Bigr) & \leq 4 \frac{\Var_{f_0}\bigl[\wh{T}\bigr]}{\mathbb{E}[f^2(X)]^2} \\
+so that $E[\wh{T}] - \wh{t}_b \geq E[f_0^2(X)]/2$. Hence, by Chebyshev's inequality
+\begin{align*}
+\mathbb{P}_{f_0}\Bigl(\wh{T} \leq \wh{t}_b \Bigr) & \leq 4 \frac{\Var_{f_0}\bigl[\wh{T}\bigr]}{\mathbb{E}[f_0^2(X)]^2} \\
 & \leq \frac{4}{n} \cdot \frac{ 3 + p_{\max}\bigl(M^4 + \|f_0\|_{\Leb^2(\Xset)}^2 \bigr)}{p_{\min}^2 \|f_0\|_{\Leb^2(\Xset)}^4} \\
-& \leq \frac{1}{16b^2}\Bigl(3 + p_{\max}n^{-1/2} + p_{\max}M^4\Bigr).
+& \leq \frac{1}{16b^2}\Bigl(3 + \frac{4bp_{\max}}{p_{\min}n^{1/2}} + p_{\max}M^4\Bigr).
 \end{align*}
@@
 \subsection{Two convenient estimates}
-The following Lemma provides convenient upper and lower bounds on our estimation variance term.
+\label{subsec:convenient_estimate}
+The following Lemmas provides convenient upper and lower bounds on our estimation variance term (Lemma~\ref{lem:variance_term_estimation}) and testing variance term (Lemma~\ref{lem:variance_term_testing}).
 \begin{lemma}
@@
 	\begin{equation*}
-	\min\Bigl\{\frac{1}{4}\floor{t^{-d/2}},n\Bigr\} \leq \sum_{k = 1}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^2 \leq t^{-d/2} +
+	\frac{1}{4}\min\Bigl\{\floor{t^{-d/2}} - 1 \vee 0,n\Bigr\} \leq \sum_{k = 2}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^2 \leq t^{-d/2} +
 	\begin{cases*}
@@
 	\begin{equation*}
-	\min\Bigl\{\frac{1}{16}\floor{t^{-d/2}},n\Bigr\} \leq \sum_{k = 1}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^4 \leq 2t^{-d/2}.
+	\frac{1}{16}\min\Bigl\{\floor{t^{-d/2}} - 1 \vee 0,n\Bigr\} \leq \sum_{k = 2}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^4 \leq 2t^{-d/2}.
 	\end{equation*}
@@
 \end{align*}
-and the various upper bounds (for $d < 4$, $d = 4$, and $d > 4$) follow by doing a straightforward integral.
-
-For the lower bound, we simply recognize that when $k \leq \floor{t^{-d/2}}$,
-\begin{equation*}
-\biggl(\frac{1}{tk^{2/d} + 1}\biggr)^2 \geq \frac{1}{4}.
-\end{equation*}
+and the various upper bounds (for $d < 4$, $d = 4$, and $d > 4$) follow upon computing the integral.
+
+For the lower bound, we simply recognize that for each $k = 2,\ldots,n$ such that $k \leq \floor{t^{-d/2}}$, it holds that $1/(tk^{2/d} + 1)^2 \geq 1/4$, and there are at least $\min\Bigl\{\floor{t^{-d/2}} - 1 \vee 0,n\Bigr\}$ such values of $k$.
 
@@
 	\begin{equation*}
-	\Pbb\Bigl[U - N \geq 2\sqrt{N t} + 2t\Bigr] \leq \exp(-t).
+	\Pbb\Bigl[U\geq 2\sqrt{N t} + 2t\Bigr] \leq \exp(-t).
 	\end{equation*}

### 2021-02-14 manuscripts/laplacian_smoothing/arXiv/appendix.tex
arXiv version updated to reflect latest version of Laplacian Smoothing paper.

--- 2021-02-14 manuscripts/laplacian_smoothing/AISTATS/supplement_minimax_optimal_laplacian_smoothing.tex
@@
-\documentclass[twoside]{article}
-\usepackage[accepted]{aistats2021}
-\usepackage{amsmath, amsfonts, amsthm, amssymb}
-\usepackage[parfill]{parskip}
-\usepackage{algpseudocode}
-\usepackage{algorithm}
-\usepackage{enumerate}
-\usepackage[shortlabels]{enumitem}
-\usepackage{mathtools}
-\usepackage{tikz}
-\usepackage{verbatim}
-
-\usepackage{natbib}
-\renewcommand{\bibname}{REFERENCES}
-\renewcommand{\bibsection}{\subsubsection*{\bibname}}
-
-\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
-\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
-\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
-\DeclareMathAccent{\wb}{0}{mathx}{"73}
-
-\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
-\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}
-
-% Make a widecheck symbol (thanks, Stack Exchange!)
-\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
-\DeclareFontShape{U}{mathx}{m}{n}{
-	<5> <6> <7> <8> <9> <10>
-	<10.95> <12> <14.4> <17.28> <20.74> <24.88>
-	mathx10
-}{}
-\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
-\DeclareFontSubstitution{U}{mathx}{m}{n}
-\DeclareMathAccent{\widecheck}{0}{mathx}{"71}
-% widecheck made
-
-\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
-\newcommand{\Graph}{\mathcal{G}}
-\newcommand{\Reals}{\mathbb{R}}
-\newcommand{\iid}{\overset{\text{i.i.d}}{\sim}}
-\newcommand{\convprob}{\overset{p}{\to}}
-\newcommand{\convdist}{\overset{w}{\to}}
-\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
-\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
-\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
-\newcommand{\iset}{\mathbf{i}}
-\newcommand{\jset}{\mathbf{j}}
-\newcommand{\myexp}[1]{\exp \{ #1 \}}
-\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
-\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
-\newcommand{\ext}[1]{\widetilde{#1}}
-\newcommand{\set}[1]{\left\{#1\right\}}
-\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
-\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
-\newcommand{\Var}{\mathrm{Var}}
-\newcommand{\Cov}{\mathrm{Cov}}
-\newcommand{\diam}{\mathrm{diam}}
-
-\newcommand{\emC}{C_n}
-\newcommand{\emCpr}{C'_n}
-\newcommand{\emCthick}{C^{\sigma}_n}
-\newcommand{\emCprthick}{C'^{\sigma}_n}
-\newcommand{\emS}{S^{\sigma}_n}
-\newcommand{\estC}{\widehat{C}_n}
-\newcommand{\hC}{\hat{C^{\sigma}_n}}
-\newcommand{\vol}{\text{vol}}
-\newcommand{\spansp}{\mathrm{span}~}
-\newcommand{\1}{\mathbf{1}}
-
-\newcommand{\Linv}{L^{\dagger}}
-\DeclareMathOperator*{\argmin}{argmin}
-\DeclareMathOperator*{\argmax}{argmax}
-
-\newcommand{\emF}{\mathbb{F}_n}
-\newcommand{\emG}{\mathbb{G}_n}
-\newcommand{\emP}{\mathbb{P}_n}
-\newcommand{\F}{\mathcal{F}}
-\newcommand{\D}{\mathcal{D}}
-\newcommand{\R}{\mathcal{R}}
-\newcommand{\Rd}{\Reals^d}
-\newcommand{\Nbb}{\mathbb{N}}
-
-%%% Vectors
-\newcommand{\thetast}{\theta^{\star}}
-\newcommand{\betap}{\beta^{(p)}}
-\newcommand{\betaq}{\beta^{(q)}}
-\newcommand{\vardeltapq}{\varDelta^{(p,q)}}
-\newcommand{\lambdavec}{\boldsymbol{\lambda}}
-
-%%% Matrices
-\newcommand{\X}{X} % no bold
-\newcommand{\Y}{Y} % no bold
-\newcommand{\Z}{Z} % no bold
-\newcommand{\Lgrid}{L_{\grid}}
-\newcommand{\Dgrid}{D_{\grid}}
-\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}
-\newcommand{\Lap}{L}
-\newcommand{\NLap}{{\bf N}}
-\newcommand{\PLap}{{\bf P}}
-\newcommand{\Id}{I}
-
-%%% Sets and classes
-\newcommand{\Xset}{\mathcal{X}}
-\newcommand{\Vset}{\mathcal{V}}
-\newcommand{\Sset}{\mathcal{S}}
-\newcommand{\Hclass}{\mathcal{H}}
-\newcommand{\Pclass}{\mathcal{P}}
-\newcommand{\Leb}{L}
-\newcommand{\mc}[1]{\mathcal{#1}}
-
-%%% Distributions and related quantities
-\newcommand{\Pbb}{\mathbb{P}}
-\newcommand{\Ebb}{\mathbb{E}}
-\newcommand{\Qbb}{\mathbb{Q}}
-\newcommand{\Ibb}{\mathbb{I}}
-
-%%% Operators
-\newcommand{\Tadj}{T^{\star}}
-\newcommand{\dive}{\mathrm{div}}
-\newcommand{\dif}{\mathop{}\!\mathrm{d}}
-\newcommand{\gradient}{\mathcal{D}}
-\newcommand{\Hessian}{\mathcal{D}^2}
-\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
-\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}
-
-%%% Misc
-\newcommand{\grid}{\mathrm{grid}}
-\newcommand{\critr}{R_n}
-\newcommand{\dx}{\,dx}
-\newcommand{\dy}{\,dy}
-\newcommand{\dr}{\,dr}
-\newcommand{\dxpr}{\,dx'}
-\newcommand{\dypr}{\,dy'}
-\newcommand{\wt}[1]{\widetilde{#1}}
-\newcommand{\wh}[1]{\widehat{#1}}
-\newcommand{\wc}[1]{\widecheck{#1}}
-\newcommand{\ol}[1]{\overline{#1}}
-\newcommand{\spec}{\mathrm{spec}}
-\newcommand{\LE}{\mathrm{LE}}
-\newcommand{\LS}{\mathrm{LS}}
-\newcommand{\SM}{\mathrm{SM}}
-\newcommand{\OS}{\mathrm{OS}}
-\newcommand{\PLS}{\mathrm{PLS}}
-
-%%% Order of magnitude
-\newcommand{\soom}{\sim}
-
-%%% Theorem environments
-\newtheorem{theorem}{Theorem}
-\newtheorem{conjecture}{Conjecture}
-\newtheorem{lemma}{Lemma}
-\newtheorem{example}{Example}
-\newtheorem{corollary}{Corollary}
-\newtheorem{proposition}{Proposition}
-\newtheorem{assumption}{Assumption}
-\newtheorem{remark}{Remark}
-
-%%% So as not to overlap with theorem/lemma numbering from the main text...
-\setcounter{theorem}{4}
-\setcounter{lemma}{2}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-
-\theoremstyle{remark}
-
-%%% Reference main text
-\usepackage{xr-hyper}
-\usepackage{hyperref}
-\hypersetup{
-	colorlinks=true,
-	linkcolor=blue,
-	citecolor=blue
-}
-\externaldocument{minimax_optimal_laplacian_smoothing}
-
-%%% New equation numbering
-\renewcommand{\theequation}{A.\arabic{equation}}
-
-%%% New section numbering
-\renewcommand{\thesection}{A\arabic{section}}
-
-\begin{document}
-
-\onecolumn	
-\aistatstitle{Supplementary Materials for ``Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs''}
-
-In this supplement, we provide proofs of all results found in the paper ``Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs''.
+\noindent In this appendix, we provide proofs of all results found in the main text.
 Our main theorems (Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing_manifold}) all follow the same general proof strategy, centered on conditioning.
@@
 In Section~\ref{sec:concentration}, we state a few concentration bounds that we use repeatedly in our proofs.
-
-\vfill
-
-\clearpage
 
@@
 taking $\delta = 1$ gives the claimed upper bound.
-
-\clearpage
-
-\bibliographystyle{plainnat}
-\bibliography{../../../graph_regression_bibliography}
-
-\end{document}

### 2021-02-14 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
arXiv version updated to reflect latest version of Laplacian Smoothing paper.

--- 2021-02-14 manuscripts/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
@@
 \documentclass[twoside]{article}
-\usepackage[accepted]{aistats2021}
 \usepackage{amsmath, amsfonts, amsthm, amssymb}
@@
 
-\runningtitle{Minimax Optimal Laplacian Smoothing}
+\begin{center} {\Large{\bf{Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs}}}
 
-\twocolumn[
+	\vspace*{.3cm}
 
-	\aistatstitle{Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs}
+	{\large{
+			\begin{center}
+				Alden Green~~~~~ Sivaraman Balakrishnan~~~~~ Ryan J. Tibshirani\\
+				\vspace{.2cm}
+			\end{center}
 
-	\aistatsauthor{Alden Green, Sivaraman Balakrishnan, and Ryan J. Tibshirani}
+			\begin{tabular}{c}
+				Department of Statistics and Data Science \\
+				Carnegie Mellon University
+			\end{tabular}
 
-	\aistatsaddress{Carnegie Mellon University}]
+			\vspace*{.2in}
+
+			\begin{tabular}{c}
+				\texttt{\{ajgreen,siva,ryantibs\}@stat.cmu.edu}
+			\end{tabular}
+	}}
+
+	\vspace*{.2in}
+
+	\today
+	\vspace*{.2in}
+
+\end{center}
 
@@
 We comment on~\ref{asmp:ls_kernel_radius_estimation} after stating Theorem~\ref{thm:laplacian_smoothing_estimation1}.
-The proof of this theorem, as with the proofs of all results in this paper, can be found in the supplementary document.
+The proof of this theorem, as with the proofs of all results in this paper, can be found in the appendices.
 \begin{theorem}
@@
 Assuming that the extension of $\wh{f}$ and the regression function $f_0$ are both suitably smooth, tools from empirical process theory (Chapter 14 of \citet{wainwright2019}) or approximation theory (Section 15.5 of \citet{johnstone2011}) should guarantee that the $\Leb^2(\Xset)$ error is not too much greater than its in-sample counterpart.
-In fact, as we show in our supplementary material, if $f_0$ is Lipschitz smooth and we extend $\wh{f}$ to be piecewise constant over the Voronoi tesselation induced by $X_1,\ldots,X_n$, the out-of-sample error $\|f_0 - \wh{f}\|_{\Leb^2(X)}$ is within a negligible factor of the in-sample error $\|f_0 - \wh{f}\|_n$, but we leave analysis of the Sobolev case to future work.
+In fact, as we show in Section~\ref{subsec:laplacian_smoothing_estimation1_pf}, if $f_0$ is Lipschitz smooth and we extend $\wh{f}$ to be piecewise constant over the Voronoi tesselation induced by $X_1,\ldots,X_n$, the out-of-sample error $\|f_0 - \wh{f}\|_{\Leb^2(X)}$ is within a negligible factor of the in-sample error $\|f_0 - \wh{f}\|_n$, but we leave analysis of the Sobolev case to future work.
 	\item Under the stronger assumption that $f_0$ is Lipschitz smooth, we can replace the factor of $\delta$ in the high probability bound by the sharper $\delta^2/n$.
@@
 	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_2d.pdf}
-	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_3d.pdf}\\ 
+	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_3d.pdf}
 	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_4d.pdf}
@@
 (Clearly the test $\wh{\varphi}$ depends on $b$, but we suppress this notationally.) The number $b \geq 1$ is a tuning parameter selected by the user, with the choice made based on the tolerated level of Type I and Type II error.
-In fact, the threshold $\wh{t}_b$ is precisely the right choice to control the Type I error of $\wh{\varphi}$; as we show in our supplementary material,
+In fact, the threshold $\wh{t}_b$ is precisely the right choice to control the Type I error of $\wh{\varphi}$; as we show in Section~\ref{sec:fixed_graph_error_bounds},
 \begin{equation}
@@
 \bibliography{../../../graph_regression_bibliography} 
+\clearpage
+
+\appendix
+
+\input{appendix.tex}
 

### 2021-02-17 notes/misc/insights_and_connections.tex
Updated the insights and connections document.

--- 2021-02-11 notes/reference_notes/insights_and_connections.tex
@@
 We use \emph{Laplacian learning} as a catch-all term for learning algorithms which leverage (possibly discrete approximations of) the Laplace operator to conduct \{unsupervised, semi-supervised, supervised\} learning tasks.
-In this note, I will detail various physical analogies and intuitions to explain what Laplacian learning is ``doing'', draw connections between Laplacian learning and other types of learning (e.g.
-kernel smoothing, RKHS regression), and hopefully provide some insight regarding when and why Laplacian learning is a reasonable idea.
-Much of what follows will be review of others' ideas, and I provide references where appropriate.
+The purpose of this note is to explain what Laplacian learning does, and when and why it works well.
+The nature of these explanations differ, based on whether the problem at hand falls into one of two regimes--- either the \emph{well-posed} or \emph{ill-posed} regime--- as I now explain.
 
-\section{Physical analogies and intuitions} 
-One theme of the literature regarding Laplacian learning is that algorithms which use the Laplace operator are sometimes well-posed, and sometimes not.
-We will be particularly interested in analogies and intuition that explain when and why algorithms that use the Laplace operator succeed, and when and why they fail.
-The question of well-posedness often hinges on the extent to which constraints enforced on some subset of an ambient domain $\Omega$ ensure a well-posed problem: that is, a problem with a reasonable solution on $\Omega$.
-We will call the constraints \emph{boundary conditions}.
+There is a long line of literature addressing the relationship between discrete Laplacian learning methods and their continuum counterparts.
+This literature provides compelling evidence that graph Laplacian learning methods are often approximating a reasonable, and indeed quite interesting, continuum counterpart.
+I term this setting the \emph{well-posed} regime.
+In this regime, I view graph Laplacians as a noisy version of the continuum problem we wish to solve, and I think the question ``what are graph Laplacians doing'' is always best answered by investigating the structure of the continuum counterpart.
+Of course, this leads to the question ``why use graph Laplacians at all''?
+The answer to this is that even though we are interested in solving these continuum problems, they are typically challenging to directly solve, which justifies use of a discrete approximation of the problem, i.e.
+graphs.
 
+On the other hand, another theme in the literature on Laplacian learning is that Laplacian learning problems are often \emph{ill-posed}---meaning, in a rough sense, that the solutions to these problems are either undesirable or even nonsensical.
+On graphs, this theme manifests itself in one of two ways: either graph Laplacian learning methods (a) are not well-posed, and do not correspond with well-posed continuum counterparts, or even more strikingly (b) are well-posed even when their seemingly obvious continuum counterparts are not.
+In this regime, in order to answer ``what are graph Laplacians doing?'' it feels like we must first tackle this question of well and ill-posedness.
+
+Because I find the first regime to be (relatively) more straightforward, \textcolor{red}{for now} in this note I focus on the second regime, paying special attention to the differences between when we fall in situation (a) vs.
+when we fall in situation (b).
+I detail various physical analogies and intuitions to explain ``what the heck Laplacian learning is doing''\footnote{Ryan's quote!}, both when it is succeeding (a) and failing (b).
+I also draw connections between Laplacian learning and other types of learning (e.g.
+RKHS regression, kernel smoothing).
+Throughout, I adopt the view that although the original question was ``what is Laplacian learning doing?'', this is simply an intermediate question along the road to ``when and why should I use Laplacian learning?'' Finally: much of what follows will be review of others' ideas, and I provide references where appropriate.
+
+\textcolor{red}{(TODO): Obviously when framed this way, it seems important to give a good summary of the well-posed setting too.}
+
+\section{Physical Analogies and Intuitions} 
+\label{sec:analogies_and_intution}
+
+We will pay special attention to the \emph{semi-supervised learning} task, because it provides a useful intermediary between discrete and continuum problems.
 Suppose we observe data $(x_1,y_1),\ldots,(x_n,y_n)$ and additionally $x_{n + 1},\ldots,x_{N}$, with $\Omega_N = \{x_1,\ldots,x_N\} \subset \Rd$.
+Typically, although not always, we will treat the samples $x_1,\ldots,x_N$ as independent samples from a distribution $P$ with support $\Omega \subseteq \Rd$, and the labels $y_i$ as being (possibly noisy) evaluations of an underlying unknown regression function $f_0$.
+Our goal is to learn the regression function $f_0$ on $x_{n + 1},\ldots,x_{N}$, using the labels $Y = \{y_1,\ldots,y_n\}$ and a Laplace operator $L_{N,\varepsilon}$, which we now define.
+
 For a given kernel $\eta(\cdot): [0,\infty) \to [0,\infty)$, and length scale (or bandwidth) $\varepsilon$, the Laplace operator $L_{N,\varepsilon}: H^2(\Omega_N) \to L^2(\Omega_N)$ is defined according to,
@@
 the connection being that $b_{N,\varepsilon}(u) = \dotp{u}{L_{N,\varepsilon}u}_{N}$ for any $u \in H^2(\Omega_N)$.
-Our goal is to learn a function $u$ on $x_{n + 1},\ldots,x_{N}$ using the labels $Y = \{y_1,\ldots,y_n\}$ and the Laplace operator $L_{N,\varepsilon}$.
 
@@
 \end{equation}
-and takes the explicit form $\wh{u} = L^{\dagger} A y$. Here $A \in \Reals^{N \times n}$ is the kernel smoothing matrix with entries $A_{ij} = \eta(\|x_i - x_j\| / \varepsilon)$.
+and takes the explicit form $\wh{u} = L_{\mathrm{un}}^{\dagger} A y$, where $L_{\mathrm{un}}$ is the Laplacian matrix over the unlabeled data only.
+Here $A \in \Reals^{N \times n}$ is the kernel smoothing matrix with entries $A_{ij} = \eta(\|x_i - x_j\| / \varepsilon)$.
 
@@
 On the other hand,~\eqref{eqn:min_energy_interpolant} provides a way to relate $\wh{u}$ to the expected hitting time.
-Concretely, supposing at the labeled points $y_i = g(x_i)$ for some Lipschitz function $g$ with Lipschitz constant $L$, then at an unlabeled point $x \in \Omega_N$ a martingale argument shows that
-\begin{equation*}
-\bigl|\wh{u}(x) - g(x)\bigr|^2 \lesssim \Ebb[\tau] \cdot \Ebb[|X_{\tau} - x|^2]  \cdot L,
-\end{equation*}
-where $\Ebb[\tau] = H\bigl(x,\{x_1,\ldots,x_n\}\bigr)$. \textcolor{red}{TODO: Calder2020a makes a lower bound argument, which also belongs here.}
+Concretely, suppose at the labeled points $y_i = g(x_i)$ for some Lipschitz function $g$ with Lipschitz constant $L$.
+Then at an unlabeled point $x \in \Omega_N$, a martingale argument shows that
+\begin{equation}
+\label{eqn:connection_hitting_time_laplace_learning}
+\bigl|\wh{u}(x) - g(x)\bigr|^2 \asymp \Ebb[\tau] \cdot \Ebb[|X_{\tau} - x|^2]  \cdot L,
+\end{equation}
+where $\Ebb[\tau] = H\bigl(x,\{x_1,\ldots,x_n\}\bigr)$.
+(In~\eqref{eqn:connection_hitting_time_laplace_learning}, the $\lesssim$ direction holds for any $L$-Lipschitz function, while the $\gtrsim$ holds for some $L$-Lipschitz function.)
 
@@
 	\item Conclude that the crucial point is the ratio of labeled to unlabeled points, as this determines whether the walk will have mixed before it arrives at a labeled point.
-	\item Discuss other fixes---three of which are ``smearing the labels''---I have not seen anyone mention this, to the best of my knowledge---changing the graph \textcolor{red}{(Shi, Calder)}, and changing the norm \textcolor{red}{(El-Alaoui, Thorpe)}.
+	\item Discuss other fixes to the ill-posedness of the semi-supervised problem when $N \to \infty$ and $n$ held constant.
+Three such fixes are (1) ``smearing the labels''---mentioned by Dejan to Ryan, but not analyzed in the literature as far as I know--- (2) changing the graph \textcolor{red}{(Shi, Calder)}, (3) changing the norm \textcolor{red}{(El-Alaoui, Thorpe)}.
 \end{itemize}
@@
 
-To be clear, by future directions I mean only things for me to track down in the future, rather than any claim of novelty.
-A tremendous amount of work has been done regarding Laplace learning and I have certainly only scratched the surface of it in this summary.
+To be clear, by future directions I mean only things for me to track down in the future, rather than any claim of novelty, although some of these ideas I have not seen proposed elsewhere.
+A tremendous amount of work has been done regarding Laplace learning and I have likely only scratched the surface of it in this summary.
 \begin{itemize}
@@
 
-\section{Laplacian learning as Reproducing Kernel Hilbert Space regression}
+\section{Laplacian learning and Reproducing Kernel Hilbert Space regression}
 We refer to the penalized analog of~\eqref{eqn:min_energy_interpolant} as \emph{Laplacian smoothing}, defined for a given value of $\lambda$ as
-\begin{equation*}
+\begin{equation}
+\label{eqn:laplacian_smoothing}
 \wt{u} := \argmin \|y - u\|_n^2 + \lambda b_r(u)
-\end{equation*}
-When the response is zero-mean, Laplacian smoothing is equivalent to ridge regression in the Reproducing Kernel Hilbert Space (RKHS) defined by the kernel $K = L^{\dagger}$.
-It is immediate that the kernel $K$ defines an RKHS $H_0^1(\Omega_N)$: $L^{\dagger}$ is a positive semi-definite (psd) bilinear form, and it is automatically trace-class because $L^2(\Omega_N)$ is finite-dimensional.
-The feature map is $\phi: x \mapsto \bigl(\lambda_2^{1/2} v_2(x),\lambda_3^{1/2} v_3(x),\ldots,\lambda_N^{1/2} v_N(x)\bigr)$.
-The norm is $\|u\|_{H_0^1(\Omega_N)} = \sqrt{b_{\varepsilon}(u)}$.
+\end{equation}
+When the response is zero-mean, Laplacian interpolation is equivalent to linear regression, and Laplacian smoothing equivalent to ridge regression, in the Reproducing Kernel Hilbert Space (RKHS) induced by the kernel $K = L^{\dagger}$.
+It is immediate that the kernel $K$ defines an RKHS: $L^{\dagger}$ is a positive semi-definite (psd) bilinear form, and it is automatically trace-class (meaning it has a discrete spectrum, with eigenvalues belonging to $\ell^1(\mathbb{N})$) because $L^2(\Omega_N)$ is finite-dimensional.
+We denote the RKHS defined by $K$ as $H_0^1(\Omega_N)$, with the subscript $0$ alluding to the fact that functions in $H_0^1(\Omega_N)$ are zero-mean.
+The feature map is $\phi: x \mapsto \bigl(\lambda_2^{-1/2} v_2(x),\lambda_3^{-1/2} v_3(x),\ldots,\lambda_N^{-1/2} v_N(x)\bigr)$.
+The (squared-)norm is $\|u\|_{H_0^1(\Omega_N)}^2 = b_{\varepsilon}(u)$.
 
@@
 When $d \geq 2$, as $N \to \infty$ and $\varepsilon \to 0$ (second condition optional), and $\lambda$ remains fixed, $\wt{u}$ will converge to the ill-posed limit~\eqref{eqn:min_energy_interpolant_illposed}.
-This can be explained by the fact that while for any finite $N$, $L^{\dagger}$ is trace-class, when $d \geq 2$ asymptotically the trace explodes, meaning
+This can be explained by the fact that---while for any finite number of samples $N$, the (pseudo-)inverse Laplacian $L^{\dagger}$ is trace-class---when $d \geq 2$ asymptotically the trace of $L^{\dagger}$ explodes, meaning
 \begin{equation*}
@@
 \end{equation*}
-This will not be true when $d = 1$.
-
 Equivalently, while it is the case that the evaluation functional $L_x: f \mapsto f(x)$ is bounded for any finite $N$,
@@
 \end{equation*}
-asymptotically the statement is not true: $\lim_{N \to \infty} C(N) = \infty$.
+asymptotically the statement is not true: $\lim_{N \to \infty, \varepsilon \to 0} C(N) = \infty$. In each case, the condition $\varepsilon \to 0$ is optional.
+
+On the other hand, when $d = 1$ and $\varepsilon \to 0$ suitably fast as $N \to \infty$, the trace converges and the evaluation functionals are uniformly bounded.
+In the limit, Laplacian smoothing converges to a ridge regression estimate in a suitable RKHS.
 
 \paragraph{Laplacian learning and RKHS regression: well-posed regime.}
-We now give a simple example showing that when $n/N \gg \varepsilon^2$, the representation cost (in $H_0^1(\Omega_N)$) of a spiky interpolant such as~\eqref{eqn:min_energy_interpolant_illposed} explodes.
-This is known to be exactly the minimum labeling rate at which minimum Dirichlet energy interpolation converges to a sensible continuum limit.
+We now give a simple example showing that when $n/N \gg \varepsilon^2$, the representation cost (in $H_0^1(\Omega_N)$) of a spiky interpolant such as~\eqref{eqn:min_energy_interpolant_illposed} explodes.\footnote{We use the phrase ``representation cost'' interchangeably with RKHS norm, and Dirichlet energy.} The rate $n/N \asymp \varepsilon^2$ is known to be the minimum well-posed labeling rate: that is, the minimum fraction of labels for which minimum Dirichlet energy interpolation converges to a sensible continuum limit, at least in the zero-noise regime.
 
@@
 \end{equation*}
-This is to be contrasted with the interpolator $u = \bigl(g(x_1),\ldots,g(x_n), g(x_{n + 1}),\ldots,g(x_N)\bigr)$, which has finite representational cost in the limit as $n \to \infty, N \to \infty$, regardless of the scaling of $n$ with respect to $N$.
+This is to be contrasted with the original regression function $g$.
+Clearly, in the zero-noise setting $g$ has zero in-sample error, the same as $u_N$.
+On the other hand, it is not hard to see that $g$ has finite representational cost in the limit as $n \to \infty, N \to \infty$, regardless of the scaling of $n$ with respect to $N$.
+As a result, we can rule out the possibility that $\wt{u} = u_N$.
+
+\subsection{Laplacian learning, RKHS regression, and continuum limits}
+\label{subsec:laplacian_learning_continuum_limits}
+In this section, we explore some connections between Laplacian learning methods and more classical notions of RKHS regression.
+Classically, RKHS regression involve minimizing loss over a function class constructed using a kernel $\mc{K}: \Omega \times \Omega \to \Reals$.
+(Note the difference between $\mc{K}$ and the kernel $K = L^{\dagger}$, a data dependent kernel with domain $\Omega_N \times \Omega_N$.)
+
+\paragraph{Laplacian learning and RKHS: continuum limits when $d = 1$.}
+Let $L_0^2(P) \subseteq \Leb^2(P)$ consist of the square-integrable functions $f$ which satisfy Neumann conditions at the boundary of $\Omega$, i.e $\frac{\partial}{\partial \mathbf{n}}f = 0$, where $\mathbf{n}$ is the normal vector.
+When $d = 1$, the continuum counterpart for the Dirichlet energy over $L_0^2(P)$ is
+\begin{equation*}
+\dotp{\Delta_{P}f}{f}_{\Leb^2(P)} = \int \|\nabla f\|^2 p^2 \,d\nu,
+\end{equation*}
+and we let $H_0^1(P) \subset L_0^2(P)$ consist of those functions $f$ with finite Dirichlet energy. Here $\Delta_P$ is the \emph{weighted Laplace-Beltrami} operator
+\begin{equation*}
+\Delta_Pf := -\frac{1}{p} \dive(p^2 \nabla f).
+\end{equation*}
+If $\Omega$ is a compact domain, and $p$ is bounded away from $0$ and $\infty$, then the Green's function of $\Delta_P$ is associated with an RKHS.
+
+\paragraph{Laplacian learning and RKHS: continuum limits when $d \geq 2$.}
+\begin{itemize}
+	\item When $d \geq 2$, the space $H_0^1(P)$ is no longer an RKHS.
+	\item I claim that the right continuum counterpart to Laplacian smoothing is instead given by restricting the domain of minimization to the first $N$ eigenfunctions of $\Delta_P$.
+In other words, let $V_k = \mathrm{span}\{v_1,\ldots,v_k\}$, and take
+	\begin{equation}
+	\label{eqn:bandlimited_laplacian_smoothing}
+	\check{u} := \argmin_{u \in V_N} \Bigl\{ \|y - u\|_n^2 + \lambda \int \|\nabla u\| p^2 \,d \nu  \Bigr\}
+	\end{equation}
+	The vector space $V_k$, equipped with inner product $\dotp{\cdot}{\cdot}_{\mc{H}}$, is an RKHS.
+Indeed, it is well-studied in the special case $d = 1$, $\Omega = \Reals$ and $P$ uniform.
+I further claim that as long as $N \to \infty$ sufficiently slowly as $n \to \infty$, the solution to~\eqref{eqn:bandlimited_laplacian_smoothing} is well-behaved in the limit.
+	\item \textcolor{red}{(TODO):} We need to justify each of the claims in the prior point. 
+	\begin{itemize}
+		\item I believe the first claim can be rigorously justified when the domain is $[0,1]^d$, the design points are equi-spaced grid points, the graph is the lattice, and the measure $P$ is uniform.
+It is much trickier to justify when the graph is a neighborhood graph, because it involves making statements about the top eigenvectors of the Laplacian $L_{N,\varepsilon}$, which are poorly understood.
+If we cannot rigorously justify it in this case, then the next best option is to give some intuition.
+For instance, we could say that all eigenvectors of $L_{N,\varepsilon}$ can be extrapolated to locally Lipschitz functions on $\Omega$ with bounded Lipschitz constant, and that such functions in turn lie very close to $V_N$.
+		\item The second claim is simply a matter of adapting some standard results regarding the entropy of RKHS.
+	\end{itemize}
+	\item We should also give some insight into the structure of $V_N$.
+For instance, what are the smoothness properties of functions $u \in V_N$?
+What does the kernel function look like (this is well known, I believe, when $d = 1$ and $p$ is uniform)?
+How do these depend on the density $p$?
+\end{itemize}
+
+\subsection{Laplacian Learning and Over-parameterized Linear Regression}
+Recent works (Belkin19, Hastie19, Bartlett19, Muthukumar19) have analyzed the behavior of minimum-norm linear regression in the noisy over-parameterized regime, where the number of features $N$ is greater than the number of samples $n$, and the responses $Y$ are assumed to be noisy evaluations of the regression function $f_0$.
+The minimum-energy interpolant $\wh{u}$ can be easily recast as the minimum-norm solution to an under-determined linear regression problem, as follows: letting
+\begin{equation}
+\begin{aligned}
+& \wh{\beta} = \argmin \|\beta\|_2^2, \\
+~~\textrm{such that} ~~& K^{1/2}\beta(x_i) = y_i~~\textrm{for~$i = 1,\ldots,n$.}
+\end{aligned}
+\end{equation}
+then $\wh{u} = K^{1/2} \wh{\beta}$.
+
+ Many of the phenomena well-known in Laplacian based semi-supervised learning can be re-expressed in the language of over-parameterized linear regression.
+In the following, let us assume that $d \geq 2$ and $\varepsilon \asymp (\log N/N)^{1/d}$.
+\begin{itemize}
+	\item The ill-posed regime for Laplacian learning is the regime where the risk converges to the risk of a \emph{null model}.
+It is the regime where the decay of the eigenvalues of $L^{\dagger}$ is too gradual, or equivalently the regime where the covariance matrix $\Sigma$ of the feature map $\phi$ is too close to isotropic.
+	\item Fix the number of labeled points $n$, and vary the number of unlabeled points $N - n$ from $0$ to $\infty$.
+I claim that the risk curve of $\wh{u}$, as a function of $N$, will replicate a descent of a \emph{double descent} curve.
+This prediction is based on the following three observations.
+	\begin{itemize}
+		\item As $N \to \infty$, the risk of $\wh{u}$ will converge to the null risk, as just identified.
+		\item As $N \to 0$, the risk of $\wh{u}$ will also converge to a constant.
+For instance, when $N = 1$, we know that $\wh{u} = Ay$.
+In words, the minimum Dirichlet energy interpolant is just performing kernel smoothing, and the choice of bandwidth results in a variance of constant order.
+		\item When $N \asymp n/\varepsilon^2$, as already mentioned the estimator $\wh{u}$ is consistent in the zero-noise setting.
+Of course, this point does not imply that $\wh{u}$ is consistent when the responses are observed with noise, but I do find it suggestive.
+	\end{itemize} 
+	\item A particular instantiation of the \emph{semi-supervised} or \emph{cluster} assumption---in which the regression function $g$ is assumed to ``align'' with the density $p$ from which the design points $\Omega_N$ were sampled---can be formulated in terms of alignment between the regression function $f_0$ and the leading eigenvectors of $\Sigma$.
+It is known that the alignment between regression function $f_0$ and the leading eigenvectors of $\Sigma$ plays a key role in how much regularization and over-parameterization should be used.
+These results can be translated into making precise statements about how much unlabeled data is ``helping'' in the semi-supervised problem.
+Making such precise statements has proved surprisingly elusive.
+\end{itemize}
+Two nice things: 
+\begin{itemize}
+	\item Interpolation by minimizing continuum Dirichlet energy, or indeed the analogous higher-order energy,
+	\begin{equation*}
+	D_{2,k}(u) := \dotp{\Delta_P^ku}{u}_{P}
+	\end{equation*}
+	is something of a dead end in the noisy response regime.  The are two cases:
+	\begin{itemize}
+		\item When the dimension $d$ of $\Omega$ is greater than $2k$, then the problem is ill-posed, and the resulting interpolator will be spiky.
+		\item When the dimension $d$ is less than $2k$, the interpolator will be H\"{o}lder smooth with a bounded Holder norm, by the Sobolev embedding theorem (with Holder index $\alpha = k - d/2$ assuming $d$ is odd), but I claim it will still have bad test error.
+One can argue that any Holder smooth interpolator must be close to the labels $Y$ over a non-negligible portion of the unlabeled points, which is equivalent to saying that it must be far from the regression function $f_0$ at a non-negligible portion of the unlabeled points.
+((Rakhlin 2019) give a formal analysis along these lines, studying minimum RKHS norm interpolation where the RKHS in question corresponds to the Laplace kernel, and thus possesses a Sobolev-like norm.)
+	\end{itemize}
+	This is disappointing: the minimum continuum Dirichlet energy interpolant seems at first blush like a very natural object of study.
+Luckily, discrete Laplacian interpolation comes to the rescue!
+It approximates the continuum Dirichlet energy, but avoids its pathological behavior.
+	\item Viewing Laplacian learning as over-parameterized linear regression makes it far more credible that the discrete Laplacian \{interpolator,smoother\} should sometimes outperform their natural continuum analogues.
+If we think about both of the discrete and continuum problems as performing linear regression on a particular feature map---where in the discrete the feature map $\phi$ maps into $\Reals^n$, and for the continuum problem maps into $\ell^2(\mathbb{N})$---then it is much less obvious that we should prefer the ``limiting'' feature map, where the number of features is infinite.
+\end{itemize}
+
+\paragraph{Future Directions.}
+It is possible that the above is a straightforward consequence of (or at least analogous to) things that are already known, and hence is not worth pursuing further.
+In that case, I think this discussion simply serves as a nice alternative perspective---elucidated by looking at things in the frequency domain---on when and why Laplacian learning is succeeding vs.
+failing (which was the original intention!)
+
+On the other hand, if we find it interesting to pursue further, some directions:
+\begin{itemize}
+	\item When the domain is the unit cube $[0,1]^d$, the design points are equi-spaced grid points, and the graph is the lattice, all the eigenvectors and eigenvalues of $L$ are explicitly known.
+This seems like a natural warmup problem, in which should not be too challenging to compute out-of-sample error.
+	\item Empirically verify that in the noisy label setting when, say, $d = 2$, the risk of $\wh{u}$ does in fact show a U-shape.
+\end{itemize}
+
+\section{Laplacian learning as kernel smoothing}
+\textcolor{red}{(TODO)}: Sketch the problem.
 
@@
 \begin{itemize}
-	\item Relate the minimum Dirichlet interpolant to overparameterized minimum-norm linear regression. 
+	\item The equivalent kernel problem: relate the Laplacian smoother $\wt{u}$ to a kernel smoother. I see two approaches to this problem.
+	\begin{itemize}
+		\item One approach is to find the right continuum counterpart to~\eqref{eqn:laplacian_smoothing} (see Section~\ref{subsec:laplacian_learning_continuum_limits} for some ideas in this direction), and proceed to find an equivalent kernel for this continuum counterpart.
+This approach is promising insofar as finding the equivalent kernel for continuum problems (that is, problems in which the regularizer is a continuum Dirichlet energy) is well-studied in the literature \textcolor{red}{(references)}.
+It is challenging for two reasons: first, we do not always (or even usually) know the continuum counterpart for Laplacian smoothing, and second, even when we do know the continuum counterpart, it is not quite the same as those continuum problems studied in the literature---in particular, the density dependence is different.
+		\item The other approach is to use the random walk connections of Section~\ref{subsec:laplacian_learning_continuum_limits}.
+	\end{itemize}
+	\item Does the Laplacian interpolant $\wh{u}$ admit an (interesting) equivalent kernel?
+Note that this question, when posed with respect to continuum Laplacian interpolation,  can be answered in the negative---as discussed above, it is always either ill-posed or inconsistent.
 \end{itemize}
 
-\section{Laplacian learning as kernel smoothing}
-
 \end{document}

### 2021-02-17 notes/weekly_notes/12.24.20.NotesForWeek.tex
Updated the document regarding extension of Laplacian estimators.

@@
 The estimate we will consider is a kernel smoother passed over $\wh{\theta}$,
-\begin{equation*}
+\begin{equation}
+\label{eqn:kernel_smoother_laplacian_eigenmaps}
 \wh{f}(x) := T_{\varepsilon,n}\wh{\theta}_{1:n}(x).
-\end{equation*}
+\end{equation}
 In order to relate the out-of-sample error $\|\wh{f} - f^{\ast}\|_P$ to the in-sample error $\|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n$, we will assume the following.

### 2021-02-17 manuscripts/laplacian_eigenmaps/outline.tex
Updated our draft of Laplacian eigenmaps paper.

@@
 \begin{itemize}
+	\item \textcolor{red}{Belkin, Slepcev} show that for a fixed $k$, the eigenvector $v_k$ converges to the eigenfunction $u_k$, in the sense that \textcolor{red}{...}.
+\textcolor{red}{Shi, Garcia Trillos, Calder, Cheng} build on this, giving finite sample bounds, rates of convergence, and making statements uniform over $k$.
+This gives helpful intuition for what Laplacian eigenmaps is doing, i.e.
+that it is in fact approximating the \textcolor{red}{following} classical spectral series estimator.
+However, we emphasize that the minimax optimality of $\wh{f}$ or a test based on $\wh{T}$ does not follow straightforwardly from such results.
+We explain why in more detail in Section~\ref{sec:analysis}.
 	\item \cite{lee2016} considers a similar estimator --- the notable difference being the normalization of the Laplacian $L$ --- but get suboptimal rates for the supervised problem. 

### 2021-02-23 notes/misc/insights_and_connections.tex
Updated the note on insights and connections.

@@
 There is a long line of literature addressing the relationship between discrete Laplacian learning methods and their continuum counterparts.
-This literature provides compelling evidence that graph Laplacian learning methods are often approximating a reasonable, and indeed quite interesting, continuum counterpart.
+This literature provides compelling evidence that graph Laplacian learning methods are often approximating reasonable, and indeed quite interesting, continuum counterparts.
 I term this setting the \emph{well-posed} regime.
-In this regime, I view graph Laplacians as a noisy version of the continuum problem we wish to solve, and I think the question ``what are graph Laplacians doing'' is always best answered by investigating the structure of the continuum counterpart.
-Of course, this leads to the question ``why use graph Laplacians at all''?
-The answer to this is that even though we are interested in solving these continuum problems, they are typically challenging to directly solve, which justifies use of a discrete approximation of the problem, i.e.
-graphs.
+In this regime, I think the question ``what is a graph Laplacian method doing'' is always best answered by investigating the structure of the counterpart continuum method---these continuum problems are the problems we are really interested in solving.\footnote{Of course, this leads to the question ``why use graph Laplacians at all''?
+The answer to this is that even if we are always interested in solving a continuum problem, they are typically challenging to directly solve.
+On the other hand, there are many methods for efficient ways to solve problems on graphs.}
 
-On the other hand, another theme in the literature on Laplacian learning is that Laplacian learning problems are often \emph{ill-posed}---meaning, in a rough sense, that the solutions to these problems are either undesirable or even nonsensical.
+On the other hand, another theme in the literature is that Laplacian learning problems are often \emph{ill-posed}---meaning, in a rough sense, that the solutions to these problems are either undesirable or even nonsensical.
 On graphs, this theme manifests itself in one of two ways: either graph Laplacian learning methods (a) are not well-posed, and do not correspond with well-posed continuum counterparts, or even more strikingly (b) are well-posed even when their seemingly obvious continuum counterparts are not.
-In this regime, in order to answer ``what are graph Laplacians doing?'' it feels like we must first tackle this question of well and ill-posedness.
+In this regime, and in particular in setting (b), we cannot explain "what graph Laplacians are doing" by appealing to their natural continuum counterparts.
+In order to answer ``what are graph Laplacians doing?'' we must first tackle this question of well and ill-posedness.
 
-Because I find the first regime to be (relatively) more straightforward, \textcolor{red}{for now} in this note I focus on the second regime, paying special attention to the differences between when we fall in situation (a) vs.
-when we fall in situation (b).
-I detail various physical analogies and intuitions to explain ``what the heck Laplacian learning is doing''\footnote{Ryan's quote!}, both when it is succeeding (a) and failing (b).
-I also draw connections between Laplacian learning and other types of learning (e.g.
-RKHS regression, kernel smoothing).
-Throughout, I adopt the view that although the original question was ``what is Laplacian learning doing?'', this is simply an intermediate question along the road to ``when and why should I use Laplacian learning?'' Finally: much of what follows will be review of others' ideas, and I provide references where appropriate.
+Because I find the first regime to be (relatively) more straightforward, \textcolor{red}{for now} in this note I focus on the second regime, paying special attention to the differences between situation (a) vs.
+situation (b).
+I start by giving some connections meant to explain ``what the heck Laplacian learning is doing''\footnote{Ryan's quote!}, both when it is succeeding (a) and failing (b).
+Its worth repeating that making these connections is much trickier in the ill-posed regime than in the well-posed regime.
+I also draw connections between Laplacian learning and other types of learning, notably RKHS/linear regression, and kernel smoothing.
+Much of what follows will be review of others' ideas, and I provide references where appropriate.
 
-\textcolor{red}{(TODO): Obviously when framed this way, it seems important to give a good summary of the well-posed setting too.}
+\textcolor{red}{(TODO)}
+\begin{itemize}
+	\item I now realize that this note is incomplete without a comprehensive consideration of the well-posed regime, in each section. I will add this consideration.
+	\item This note includes extensive discussion of minimum energy interpolation, a minimum amount of discussion of Laplacian smoothing, and no discussion at all of Laplacian eigenmaps.
+This is because (1) there is more discussion of minimum energy interpolation than smoothing in the literature, and (2) eigenmaps have a fairly well-understood continuum limit.
+Of course, my novel contributions regarding supervised learning have been entirely related to smoothing and eigenmaps, so if anything the discussion should focus on those two methods.
+I will correct this.
+\end{itemize}
 
-\section{Physical Analogies and Intuitions} 
-\label{sec:analogies_and_intution}
+\section{Setup and Methods}
 
-We will pay special attention to the \emph{semi-supervised learning} task, because it provides a useful intermediary between discrete and continuum problems.
-Suppose we observe data $(x_1,y_1),\ldots,(x_n,y_n)$ and additionally $x_{n + 1},\ldots,x_{N}$, with $\Omega_N = \{x_1,\ldots,x_N\} \subset \Rd$.
+We define the various learning setups in which we subsequently operate. We then define some Laplace learning methods which work in these setups.
+
+\subsection{Learning setups}
+We focus on two learning setups: supervised learning and semi-supervised learning.
+
+\paragraph{Supervised learning.}
+\textcolor{red}{(TODO)}
+
+\paragraph{Semi-supervised learning.}
+Although our theoretical contributions focus on the supervised task, in this note we pay special attention to the \emph{semi-supervised learning} task.
+This is because it provides a useful intermediary between the standard Laplace-based methods for semi-supervised learning are a useful intermediary between discrete and continuum problems.
+
+Suppose we observe labeled data $(x_1,y_1),\ldots,(x_n,y_n)$ and additionally unlabeled design points $x_{n + 1},\ldots,x_{N}$, with $\Omega_N = \{x_1,\ldots,x_N\} \subset \Rd$.
 Typically, although not always, we will treat the samples $x_1,\ldots,x_N$ as independent samples from a distribution $P$ with support $\Omega \subseteq \Rd$, and the labels $y_i$ as being (possibly noisy) evaluations of an underlying unknown regression function $f_0$.
-Our goal is to learn the regression function $f_0$ on $x_{n + 1},\ldots,x_{N}$, using the labels $Y = \{y_1,\ldots,y_n\}$ and a Laplace operator $L_{N,\varepsilon}$, which we now define.
+Our goal is to learn the regression function $f_0$ on $x_{n + 1},\ldots,x_{N}$, using the labels $Y = \{y_1,\ldots,y_n\}$.
 
-For a given kernel $\eta(\cdot): [0,\infty) \to [0,\infty)$, and length scale (or bandwidth) $\varepsilon$, the Laplace operator $L_{N,\varepsilon}: H^2(\Omega_N) \to L^2(\Omega_N)$ is defined according to,
+\subsection{Methods}
+We define three methods to learn $f_0$: Laplacian interpolation, Laplacian smoothing, and Laplacian eigenmaps.
+As there names suggest, each of which them are defined with respect to the graph Laplacian.
+For a given kernel $\eta(\cdot): [0,\infty) \to [0,\infty)$, and length scale (or bandwidth) $\varepsilon$, the graph Laplacian $L_{N,\varepsilon}$ is a difference operator, acting on a function $u \in L^2(\Omega_N)$ as follows,
 \begin{equation*}
-L_{N,\varepsilon}u(x) = \frac{1}{N\epsilon^{d + 2}}\sum_{j = 1}^{N} \bigl(u(x) - u(x_j)\bigr)\eta\biggl(\frac{|x - x_i|}{\varepsilon}\biggr).
+L_{N,\varepsilon}u(x) := \frac{1}{N\varepsilon^{d + 2}}\sum_{j = 1}^{N} \bigl(u(x) - u(x_j)\bigr)\eta\biggl(\frac{|x - x_i|}{\varepsilon}\biggr).
 \end{equation*}
-The operator $L_{N,\varepsilon}$ is associated with a \emph{Dirichlet energy} $b_{N,\varepsilon}: H^1(\Omega_N) \to [0,\infty)$, defined as 
+The graph Laplacian is associated with a \emph{Dirichlet energy} $b_{N,\varepsilon}$, defined as 
 \begin{equation*}
-b_{N,\varepsilon}(u) = \sum_{i,j = 1}^{N} \bigl(u(x_i) - u(x_j)\bigr)^2 \eta\biggl(\frac{|x - x_i|}{\varepsilon}\biggr),
+b_{N,\varepsilon}(u) := \sum_{i,j = 1}^{N} \bigl(u(x_i) - u(x_j)\bigr)^2 \eta\biggl(\frac{|x - x_i|}{\varepsilon}\biggr).
 \end{equation*}
-the connection being that $b_{N,\varepsilon}(u) = \dotp{u}{L_{N,\varepsilon}u}_{N}$ for any $u \in H^2(\Omega_N)$.
+The connection is that $b_{N,\varepsilon}(u) = \dotp{u}{L_{N,\varepsilon}u}_{N}$.
 
-\paragraph{Mininum energy interpolation, and a connection with random walks.}
-
-A natural Laplace learning algorithm for semi-supervised learning is the minimum (Dirichlet) energy interpolant of $y_1,\ldots,y_n$, originally considered in the semi-supervised learning literature by \textcolor{red}{Zhu}, (and in the supervised learning case by \textcolor{red}{Duchon}) and defined as
+\paragraph{Laplacian interpolation.}
+One way to use the Laplacian in semi-supervised learning is to compute the minimum (Dirichlet) energy interpolant $\wh{u}$.
+By interpolant, we mean that $\wh{u}$ is exactly equally to the response at every labeled points, $\wh{u}(x_i) = y_i$ for $i = 1,\ldots,n$.
+The minimum energy interpolant was originally considered in the semi-supervised learning literature by \textcolor{blue}{Zhu}.
+It is quite close to thin-plate splines, suggested originally by \textcolor{blue}{Duchon}.
+Formally, $\wh{u}$ is defined as
 \begin{equation}
@@
 \end{equation}
-Some basic algebra shows that ${u}$ satisfies the \emph{Laplace equation with Dirichlet boundary conditions}, i.e. 
+
+\paragraph{Laplacian smoothing.}
+Traditionally, Laplacian interpolation is thought to be appropriate only in the zero-noise regime, when the response $y_i = f_0(x_i)$.
+In the noisy regime, where $y_i = f_0(x_i) + \varepsilon_i$, it is more standard to trade off data-fidelity against the complexity of the fit.
+One way to do this is by penalization.
+We refer to the penalized analog of~\eqref{eqn:min_energy_interpolant} as \emph{Laplacian smoothing}, which defined for a given value of $\lambda > 0$ as
+\begin{equation}
+\label{eqn:laplacian_smoothing}
+\wt{u} := \argmin \|y - u\|_n^2 + \lambda b_r(u).
+\end{equation}
+
+\paragraph{Laplacian eigenmaps.}
+\textcolor{red}{(TODO)}
+
+\section{\textcolor{red}{Laplace learning methods and their associated metrics}} 
+\label{sec:analogies_and_intution}
+
+\textbf{Goal}: The premise of this section is that Laplace learning methods all make implicit use of a non-trivial metric defined over the design points $\Omega_N$.
+The goal of this section is (a) to elucidate what this metric is, and (b) to better understand how Laplace learning methods make use of this metric.
+
+\paragraph{A connection with random walks.}
+
+Some basic algebra shows that $\wh{u}$ satisfies the \emph{Laplace equation with Dirichlet boundary conditions}, i.e. 
 \begin{equation}
@@
 \end{equation}
-and takes the explicit form $\wh{u} = L_{\mathrm{un}}^{\dagger} A y$, where $L_{\mathrm{un}}$ is the Laplacian matrix over the unlabeled data only.
+It takes the explicit form $\wh{u} = L_{\mathrm{un}}^{\dagger} A y$, where $L_{\mathrm{un}}$ is the Laplacian matrix over the unlabeled data only.
 Here $A \in \Reals^{N \times n}$ is the kernel smoothing matrix with entries $A_{ij} = \eta(\|x_i - x_j\| / \varepsilon)$.
@@
 \end{equation*}
-The resistance distance arises from two different (and therefore equivalent) physical interpretations.
-The first treats the graph $G$ corresponding to adjacency matrix $A$ as a resistor network.
+The resistance distance arises from two different physical interpretations---one in terms of resistors in electrical networks, the other in terms of random walks.
+The first interpretation treats the graph $G$ as a resistor network.
 Placing an external current $c_{ext} := e_i - e_j$ then induces a potential difference between points $x_i$ and $x_j$, and it follows from Ohm's Law and Kirchoff's Laws that this potential difference is precisely $R(i,j)$.
@@
 
-\paragraph{Connections between resistance distance and Laplace learning: ill-posed regime.}
-We know that, when dimension $d \geq 2$, both the resistance distance and the minimum-energy interpolant are ill-posed, meaning they converge to degenerate solutions as the number of unlabeled points $N \to \infty$, and the radius $\varepsilon \to 0$.
+\subsection{Well-posed regime}
+\textcolor{red}{(TODO)}: I repeat my previous observation: both of questions (a) and (b) make sense in the well-posed regime, and are worth answering carefully in that case.
+
+\subsection{Ill-posed regime, setting (a)}
+We know that, when dimension $d \geq 2$, both resistance distance and Laplace interpolation are ill-posed as the number of unlabeled points $N \to \infty$, and (optionally) the radius $\varepsilon \to 0$.
 The resistance distance, which for finite $N$ is a measure of some mixture of global and local information, in the limit contains only global information:
@@
 \section{Laplacian learning and Reproducing Kernel Hilbert Space regression}
-We refer to the penalized analog of~\eqref{eqn:min_energy_interpolant} as \emph{Laplacian smoothing}, defined for a given value of $\lambda$ as
-\begin{equation}
-\label{eqn:laplacian_smoothing}
-\wt{u} := \argmin \|y - u\|_n^2 + \lambda b_r(u)
-\end{equation}
 When the response is zero-mean, Laplacian interpolation is equivalent to linear regression, and Laplacian smoothing equivalent to ridge regression, in the Reproducing Kernel Hilbert Space (RKHS) induced by the kernel $K = L^{\dagger}$.
@@
 
-\paragraph{Laplacian smoothing as RKHS regression: ill-posed regime.}
+\paragraph{Laplacian smoothing and RKHS regression: ill-posed regime, situation (b).}
 When $d \geq 2$, as $N \to \infty$ and $\varepsilon \to 0$ (second condition optional), and $\lambda$ remains fixed, $\wt{u}$ will converge to the ill-posed limit~\eqref{eqn:min_energy_interpolant_illposed}.
@@
 
-\paragraph{Laplacian learning and RKHS regression: well-posed regime.}
-We now give a simple example showing that when $n/N \gg \varepsilon^2$, the representation cost (in $H_0^1(\Omega_N)$) of a spiky interpolant such as~\eqref{eqn:min_energy_interpolant_illposed} explodes.\footnote{We use the phrase ``representation cost'' interchangeably with RKHS norm, and Dirichlet energy.} The rate $n/N \asymp \varepsilon^2$ is known to be the minimum well-posed labeling rate: that is, the minimum fraction of labels for which minimum Dirichlet energy interpolation converges to a sensible continuum limit, at least in the zero-noise regime.
+\paragraph{Laplacian learning and RKHS regression: ill-posed regime, situation (a).}
+We now give a simple example showing that when $n/N \gg \varepsilon^2$, the representation cost (in $H_0^1(\Omega_N)$) of a spiky interpolant such as~\eqref{eqn:min_energy_interpolant_illposed} explodes.\footnote{We use the phrase ``representation cost'' interchangeably with RKHS norm, and Dirichlet energy.} In the zero-noise regime, the rate $n/N \asymp \varepsilon^2$ is known to be the minimum well-posed labeling rate: that is, the minimum fraction of labels for which minimum Dirichlet energy interpolation converges to a sensible continuum limit.
 
-Suppose $y_i = g(x_i)$ for some function $g$, and suppose (for convenience only) that $\wb{y} = 0$.
-Then $u_N = (y_1,\ldots,y_n, 0,\ldots,0)$ belongs to $H_0^1(\Omega_n)$ and has squared-norm $\|u_N\|_{H_0^1(\Omega_N)}^2 = b_r(u_N)$ of at least
+Suppose $f_0$ is a Lipschitz function and that $y_i = f_0(x_i)$.
+Also, for convenience only $\wb{y} = 0$.
+Let $u_N$ be a spiky interpolant, e.g.
+$u_N := (y_1,\ldots,y_n, 0,\ldots,0)$.
+The interpolant $u_N$ belongs to $H_0^1(\Omega_N)$ and has squared-norm $\|u_N\|_{H_0^1(\Omega_N)}^2 = b_{\varepsilon}(u_N)$ of at least
 \begin{equation*}
-b_{\varepsilon}(u_N) \geq \frac{1}{N^2 \varepsilon^{d + 2}} \sum_{i = 1}^{n} \sum_{j = n + 1}^{N} g(x_i)^2 \geq \frac{n}{N^2 \varepsilon^{d + 2}} d_{\min}(G) \|g\|_n^2.
+b_{\varepsilon}(u_N) \geq \frac{1}{N^2 \varepsilon^{d + 2}} \sum_{i = 1}^{n} \sum_{j = n + 1}^{N} \bigl(f_0(x_i)\bigr)^2 A_{ij} \geq \frac{n}{N^2 \varepsilon^{d + 2}} d_{\min}(G) \|f_0\|_n^2.
 \end{equation*} 
-In the above $d_{\min}(G) = \min\bigl\{\sum_{j = n + 1}^{N} A_{ij}: i = 1,\ldots,n\bigr\}$.
-Under standard \{random, fixed\} design conditions on $\Omega_N$, asymptotically $d_{\min}(G) \asymp N\varepsilon^d$ and $\|g\|_n^2 \to \|g\|_{\Leb^2(\Omega)} > 0$.
-Thus
+In the above $d_{\min}(G) = \min\bigl\{\sum_{j = n + 1}^{N} A_{ij}: i = 1,\ldots,n\bigr\}$, this is like the minimum degree any labeled point $x_i$ would have in a graph over the unlabeled points.
+Under standard design conditions on $\Omega_N$, asymptotically $d_{\min}(G) \asymp N\varepsilon^d$ and $\|f_0\|_n^2 \to \|f_0\|_{\Leb^2(\Omega)} > 0$.
+Thus if $n/N \gg \varepsilon^2$,
 \begin{equation*}
-\lim_{N \to \infty} b_{\varepsilon}(u_N) \geq c \cdot \lim_{N \to \infty} \frac{n}{N \varepsilon^2}.
+\lim_{N \to \infty} b_{\varepsilon}(u_N) \geq c \cdot \lim_{N \to \infty} \frac{n}{N \varepsilon^2} = \infty.
 \end{equation*}
-This is to be contrasted with the original regression function $g$.
-Clearly, in the zero-noise setting $g$ has zero in-sample error, the same as $u_N$.
-On the other hand, it is not hard to see that $g$ has finite representational cost in the limit as $n \to \infty, N \to \infty$, regardless of the scaling of $n$ with respect to $N$.
-As a result, we can rule out the possibility that $\wt{u} = u_N$.
+Now we show there are other interpolants with much smaller norm.
+For instance, the regression function itself is an interpolant, because we are in the zero-noise regime.
+Using the $L$-Lipschitz property of $f_0$, we have that
+\begin{equation*}
+b_{\varepsilon}(f_0) = \frac{1}{N^2 \varepsilon^{d + 2}} \sum_{i = 1}^{n} \sum_{j = n + 1}^{N} \bigl(f_0(x_i) - f_0(x_j)\bigr)^2 A_ij \leq L^2;
+\end{equation*}
+the second inequality again holds under standard design conditions, with high probability if the design is random.
+As a result, we can rule out the possibility that $\wt{u}$ or $\wh{u}$ are equal to $u_N$---there exists another function which interpolates the labeled data, and which has smaller norm in $H_0^1(\Omega_N)$.
 

### 2021-02-23 notes/weekly_notes/12.24.20.NotesForWeek.tex
Updated the note on insights and connections.

@@
 \begin{lemma}
-	Suppose $f^{\ast} \in H_0^{s}(\Omega)$, $p \in C^{\beta}(\Omega)$ for $\beta = s - 1$, and that $\eta$ is an order-$s$ kernel. Then
+	Suppose $f^{\ast} \in H_0^{s}(\Omega)$, $p \in C^{r}(\Omega)$ for $r = s - 1$, and that $\eta$ is an order-$s$ kernel. Then
 	\begin{equation*}
@@
 \end{lemma}
+\begin{proof}
+	\textcolor{red}{(TODO)}
+\end{proof}
 
-Thus balancing bias and variance, using a smaller bandwidth ($\varepsilon \asymp n^{-1/(2\beta + d)}$) than is typical for kernel smoothing with noisy data, leads to an asymptotically negligible error ($n^{-2s/(2\beta + d)}$) compared to the minimax rate for estimating Sobolev functions.
+Thus balancing bias and variance, using a smaller bandwidth ($\varepsilon \asymp n^{-1/(2r + d)}$) than is typical for kernel smoothing with noisy data, leads to an asymptotically negligible error ($n^{-2s/(2r + d)}$) compared to the minimax rate for estimating Sobolev functions.
 
@@
 	\begin{equation*}
-	\Bigl(\bigl(T_{n,\varepsilon}u\bigr)(x)\Bigr)^2 \leq \frac{1}{d_n(x) n} \sum_{i = 1}^{n} u_i^2 \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
+	\Bigl(\bigl(T_{n,\varepsilon}u\bigr)(x)\Bigr)^2 \leq \frac{1}{d_n(x) n} \sum_{i = 1}^{n} (u_i)^2 \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
 	\end{equation*}

### 2021-02-23 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Updated draft of Laplacian Eigenmaps manuscript. Switched to new tex file.

@@
 \usepackage{graphicx}
-\usepackage[colorlinks]{hyperref}
+\usepackage{hyperref}
+\hypersetup{
+	colorlinks=true,
+	linkcolor=blue,
+	citecolor=blue
+}
 \usepackage[parfill]{parskip}
@@
 \begin{document}
-\title{Minimax-optimal regression over Sobolev Spaces with Neighborhood Graphs}
+\title{Minimax-optimal Laplacian Eigenmaps regression over Sobolev Spaces with Neighborhood Graphs}
 \author{Alden Green}
@@
 
+\section{Introduction}
+\label{sec:introduction}
+
+\textbf{(1) Geometric graphs.} In graph-based learning, one observes data $X_1,\ldots,X_n$ sampled independently from an unknown distribution $P$, and forms a geometric graph $G$---with edges $e_{ij}$ corresponding to proximity between samples $X_i$ and $X_j$---over the observed data.
+Geometric graphs encode information about $P$ in an extremely general manner, and can thus be leveraged to conduct many different fundamental statistical tasks.
+These include clustering, semi-supervised learning, classification, and regression---both estimation and goodness-of-fit testing.
+Though much theoretical work has been done on the consistency of graph-based learning methods---and more recently, rates of convergence have been established for some problems---little is known so far about their optimality, even for classic statistical tasks.
+
+\textbf{(2) What we study: Regression and Laplacian eigenmaps.} In this paper we focus on regression where in addition to the design points $X_1,\ldots,X_n$ one observes real-valued responses $Y_1,\ldots,Y_n$, and learns the regression function $f_0(x) := E[Y|X = x]$.
+We consider both the estimation and goodness-of-fit testing problems.
+The methods we will study are based on \emph{Laplacian eigenmaps}, first introduced by \cite{belkin03a}, which projects the response vector $Y = (Y_1,\ldots,Y_n)$ onto the span of the leading eigenvectors of the graph Laplacian $L$.
+The Laplacian $L$ is a difference operator, acting on functions $f: {\bf X} \to \Reals$ as follows,
+\begin{equation}
+\label{eqn:graph_laplacian}
+Lf(X_i) = \sum_{j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)e_{ij}. 
+\end{equation}
+The Laplacian $L$ can be seen as a discretization of the Laplace-Beltrami operator $\Delta_P$ \textcolor{red}{(...)}.
+The eigenvectors of $L$ form an orthonormal basis of $L^2(X)$, and are approximations to the eigenvectors of $\Delta_P$.
+Their corresponding eigenvalues are estimates of~\textcolor{red}{(...)}, and provide a notion of smoothness for each eigenvector---the smaller the eigenvalue, the smoother the eigenvector.
+
+\textbf{(3) Why we study.} Laplacian eigenmaps is thus an example of a \emph{spectral series} estimator.
+Series estimators are one of the most classical ways to perform non-parametric regression.
+Traditionally, series estimators are defined with respect to a fixed set of basis functions, orthogonalized with respect to some pre-determined reference measure $Q$.
+In contrast, eigenvectors of the graph Laplacian are data-dependent objects, and adapt to the geometry of the distribution $P$ in a rich manner.
+
+The study of (classical) spectral series estimators is rich, and their theoretical optimality is well-understood \textcolor{red}{(references)}.
+However, in practice these estimators suffer from some serious drawbacks.
+
+\textbf{(4) Our contributions.} Laplacian eigenmaps avoid these drawbacks \textcolor{red}{(...)} On the other hand their statistical properties are not as well understood \textcolor{red}{(...)}.
+The primary contribution of our paper is to fill this theoretical gap, by answering the following question:
+
+\begin{quote}
+	\textcolor{red}{(TODO)}
+\end{quote}
+
+Broadly speaking, we show that when the regression function $f_0$ is smooth, in the sense of having bounded derivatives in the Sobolev sense, Laplacian eigenmaps is statistically minimax optimal.
+This statement holds for different relations between the dimension $d$ and number of derivatives $s$, depending on the problem (estimation) \textcolor{red}{(TODO): Add two tables summarizing your major results here.}
+
+\textbf{(5) Organization.}
+
+\textbf{(6) Notation.}
+
+\textcolor{red}{TODO}: Define $\Leb^2(\mc{X}), \Leb^2(P)$ and $\Leb^2(P_n)$.
+
+For a square, symmetric matrix $A \in \Reals^{n \times n}$, an eigenvector $v$ of $A$ is a \textcolor{red}{unit-norm} solution to the eigenproblem
+\begin{equation*}
+Av = \lambda v,
+\end{equation*}
+and $\lambda$ is the corresponding eigenvalue.
+
+\section{Setup, Background, and Overview of Results}
+\label{sec:setup_main_results}
+
+In this section, we begin by giving a precise definition of Laplacian eigenmaps.
+We then review minimax rates for non-parametric regression over Sobolev spaces, allowing us to summarize our main results.
+
+\subsection{Non-parametric regression with Laplacian Eigenmaps}
+\label{sec:regression_laplacian_eigenmaps}
+
+\textbf{(1) Formal setup.}
+We will operate in the usual setting of non-parametric regression with random design.
+We observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$, where the design points $X_1,\ldots,X_N$ are drawn i.i.d from a distribution $P$ supported on a compact set $\mc{X} \subseteq \Rd$, and responses
+\begin{equation*}
+Y_i = f_0(X_i) + \epsilon_i,
+\end{equation*}
+with regression function $f_0: \mc{X} \to \Reals$, and $\epsilon_i \sim N(0,1)$ independent Gaussian noise.
+For simplicity we will assume throughout that the noise has unit-variance, but all of our results extend in a straightforward manner to the case where the variance is equal to a known positive value.
+
+\textbf{(2) Laplacian eigenmaps.} Laplacian eigenmaps constructs an estimate of the regression function $f_0$ using the eigenvectors of a neighborhood graph Laplacian operator.
+For a positive, radially symmetric kernel $\eta: [0,\infty) \to [0,\infty)$, and a bandwidth parameter $\varepsilon > 0$, the \emph{neighborhood graph Laplacian} operator $L_{n,\varepsilon}$ is
+\begin{equation}
+\label{eqn:neighborhood_graph_laplacian}
+L_{n,\varepsilon}u(x) = \frac{1}{n\varepsilon^{d + 2}} \sum_{i = 1}^{n} \bigl(u(x) - u(X_j)\bigr) \eta\biggl(\frac{\|x - X_j\|}{\varepsilon}\biggr)
+\end{equation}
+The constant scaling $(n\varepsilon^{d + 2})^{-1}$ is purely for convenience in taking limits as $n \to \infty, \varepsilon \to 0$.
+Also, although~\eqref{eqn:neighborhood_graph_laplacian} makes sense for any $x \in \mc{X}$ and $u: \mc{X} \to \Reals$, we will typically think of $x \in \{X_1,\ldots,X_n\}$ and $u = (u(X_1),\ldots,u(X_n)) \in \Leb^2({\bf X})$.
+The graph Laplacian is positive semi-definite, and we will index its eigenpairs $(\lambda_1,v_1),\ldots,(\lambda_n,v_n)$ in ascending order of eigenvalue, $0 = \lambda_1 \leq \ldots \leq \lambda_n$.
+
+For a given $K \in \{1,\ldots,n\}$, the order-$K$ Laplacian eigenmaps estimator simply projects the response vector ${\bf Y}$ onto the first $K$ eigenvectors of $L_{n,\varepsilon}$: letting $V_K \in \Reals^{n \times K}$ be the matrix with columns $v_1,\ldots,v_K$, we have that
+\begin{equation}
+\label{eqn:laplacian_eigenmaps_estimator}
+\wh{f} := \sum_{k = 1}^{K} \dotp{Y}{v_k}_{n} v_k = \frac{1}{n} V_K V_K^{\top} Y.
+\end{equation} 
+Since the eigenvectors $v_1,\ldots,v_n$ are orthogonalized with respect to the $L^2(P_n)$ inner product, the estimator $\wh{f}$ is the least-squares solution to the linear regression problem with responses $Y_1,\ldots,Y_n$ and features $v_1,\ldots,v_K$.
+Note that $\wh{f}$ is defined only in-sample, that is, only at the design points $X_1,\ldots,X_n$---we will consider the question of out-of-sample estimation later in Section~\ref{sec:out_of_sample}.
+
+If $\wh{f}$ is a reasonable estimate of $f_0$, then the test statistic
+\begin{equation}
+\label{eqn:laplacian_eigenmaps_test}
+\wh{T} := \|\wh{f}\|_n^2
+\end{equation}
+is in turn a reasonable estimate of $\|f_0\|_{P}^2$, and can be used to distinguish whether or not $f_0 = 0$.
+
+\subsection{Sobolev Classes}
+\label{sec:sobolev}
+Before we give an overview of our main results, we will first pause to review the Sobolev classes and \textcolor{red}{associated} minimax rates.
+
+\textbf{(1) Sobolev norms, semi-norms, and balls.} 
+The Sobolev class $H^s(P)$ consists of all those functions $f \in \Leb^2(P)$ which have weak partial derivatives of all orders $j = 1,\ldots,s$, all of which are bounded in $\Leb^2(P)$ norm.
+The Sobolev norm $\|f\|_{H^s(P)}$ is defined according to
+\begin{equation*}
+\|f\|_{H^s(P)}^2 := \|f\|_{\Leb^2(P)}^2 + \sum_{j = 1}^{s} \|D^jf\|_{\Leb^2(P)}^2,
+\end{equation*}
+where $D^jf := \sum_{|\alpha| = j}D^{\alpha}f$ is the sum of all weak partial derivatives of order $j$, and we use the multiindex notation $D^{\alpha}f = \partial^j f/(\partial^{\alpha_1}x^1\ldots \partial^{\alpha_d}x^d)$ for multiindex $\alpha$.
+The Sobolev ball of radius $M$ consists of all those functions $f \in H^s(P)$ with Sobolev norm no greater than $M$,
+\begin{equation*}
+H^s(\mc{X};M) := \{f \in H^s(P): \|f\|_{H^s(P)} \leq M\}
+\end{equation*}
+
+\textcolor{red}{(TODO)}: 
+\begin{itemize}
+	\item Possibly define the iterated Laplace-Beltrami operator.
+Point out that for $H^{2s}(P)$ functions, the Sobolev semi-norm can be stated in terms of inner products using the iterated Laplace-Beltrami operator.
+	\item Introduce the notation $H^s(\mc{X}) = H^s(\nu)$, where $\nu$ is Lebesgue measure.
+\end{itemize}
+
+\textbf{(2) Zero-trace and periodic conditions.}
+When $s \geq 1$, our results regarding Laplacian eigenmaps will require some boundary assumptions on $f_0$.
+In particular, we will assume that $f_0$ belongs to the set of zero trace Sobolev functions $H_0^{s}(P)$, meaning \textcolor{red}{roughly} that $f_0 \in H^s(P)$ additionally satisfies,
+\begin{equation}
+\label{eqn:zero_trace}
+D^jf(x) = 0,~~\textrm{for all $x \in \partial \mc{X}$ and $j = 0,\ldots,s - 1$.}
+\end{equation}
+\textcolor{red}{(TODO)}: This is in spirit right but technically incoherent unless $f \in C^{\infty}(\mc{X})$. Need to use the formal definition of trace.
+
+Let us justify why we impose a boundary condition such as zero-trace.
+For this part only we concentrate on the special case of $d = 1$ and the domain $\mc{X} = [0,1]$.
+In this case, there exists another way to define Sobolev space, via a sequence space representation (equiv Fourier transform).
+Let $\psi_1,\psi_2,\ldots$ denote the trigonometric basis of $\Leb^2([0,1])$,\footnote{To be explicit, $\psi_1(x) := 1$, $\psi_{2k}(x) := \sqrt{2} \cos(2\pi k x)$ and $\psi_{2k + 1}(x) := \sqrt{2} \sin(2\pi k x)$} and let $\Theta \subset \ell^2(\mathbb{N})$ be the Sobolev ellipsoid $\Theta = \{\theta \in \ell^2(\mathbb{N}): \sum_{k = 1}^{\infty} k^{2s} \theta_k^2 \leq M\}$.
+The function class $\{f = \sum_{k = 1}^{\infty} \theta_k \psi_k: \theta \in \Theta\}$ is \textcolor{red}{equivalent to} the \emph{periodic Sobolev space} $H_{\mathrm{per}}^s([0,1])$, defined as
+\begin{equation}
+\label{eqn:periodic_sobolev_space}
+H_{\mathrm{per}}^s([0,1]) = \Bigl\{f \in H^s([0,1],M): D^jf(0) = D^jf(1)~~\textrm{for $j = 0,\ldots, s - 1$} \Bigr\}
+\end{equation}
+It is known that when $P$ is uniform, the eigenvectors $v_k$ converge to $\psi_k$ as $n \to \infty, \varepsilon \to 0$, and it is therefore natural that they accurately approximate only periodic functions $f_0 \in H^s([0,1])$.
+Although the zero-trace boundary condition is somewhat more restrictive than mere periodicity, the point is that some kind of boundary condition is inherently necessary in order to obtain sharp rates for spectral series estimators.
+For simplicity, we will stick to the specific boundary condition~\eqref{eqn:zero_trace}.
+
+\subsection{Minimax Rates}
+\label{subsec:minimax_rates_sobolev}
+We now turn to reviewing the minimax estimation and goodness-of-fit testing rates over Sobolev classes.
+We will always be interested in measuring loss in mean-squared error, that is in squared $\Leb^2$-norm.
+In this case, standard non-parametric minimax rates hold under some regularity conditions on $\mc{X}$ and $P$.
+Specifically, we will assume the following.\footnote{Assumption~\ref{asmp:sobolev_radius} ensures that we get a non-parametric rate.
+If $M \lesssim n^{-1/2}$ then both the estimation and testing rates become the parametric $n^{-1}$.}
+\begin{enumerate}[label=(A\arabic*)]
+	\item 
+	\label{asmp:domain}
+	The domain $\mc{X}$ is an open connected subset of $\Rd$, with Lipschitz boundary.
+	\item
+	\label{asmp:density}
+	The design distribution $P$ admits a density $p$ with respect to Lebesgue measure. The density $p$ is bounded away from $0$ and $\infty$: there exists $\rho \geq 1$ such that
+	\begin{equation*}
+	0 < \frac{1}{\rho} \leq p(x) \leq \rho < \infty,~~\textrm{for all $x \in \mc{X}$}
+	\end{equation*}
+	\item
+	\label{asmp:sobolev_radius}
+	The radius $M$ of the Sobolev class satisfies $M = M(n) \gtrsim n^{-1/2}$.
+\end{enumerate}
+Under assumptions~\ref{asmp:domain}-\ref{asmp:sobolev_radius}, the minimax estimation rate over $\mc{H} = H^s(\mc{X};M)$ or $\mc{H} = H_0^s(\mc{X};M)$) is (see e.g. \citet{tsybakov2008_book})
+\begin{equation}
+\label{eqn:minimax_estimation_rate}
+\inf_{\wh{f}} \sup_{f_0 \in \mc{H}} \|f_0 - f\|_P^2 \asymp M^{2d/(2s + d)}n^{-2s/(2s + d)};
+\end{equation}
+here the infimum is over estimators $\wh{f}$.
+
+In the goodness-of-fit testing problem, we ask for a test function---formally, a Borel measurable function $\phi$ that takes values in $\{0,1\}$--- which can distinguish between the hypotheses
+\begin{equation}
+\mathbf{H}_0: f_0 = f_0^{\star}, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{H} \setminus \{f_0^{\star}\}.
+\end{equation} 
+Typically, the null hypothesis $f_0 = f_0^{\star} \in \mc{F}$ reflects the absence of interesting structure, and $\mc{F} \setminus  \{f_0^{\star}\}$ is a set of smooth departures from this null.
+To fix ideas, as in \citet{ingster2009} we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\star} = 0$ and $\mc{H}$ is a Sobolev ball, $\mc{H} = H^s(\Xset,M)$ or $\mc{H} = H_0^s(\mc{X};M)$ is a Sobolev ball; this is without loss of generality.
+
+The minimax critical radius is the smallest value of $\sigma$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{H}_{\sigma} := \mc{H} \cap \{f: \|f\|_{\Leb^2(P)} \geq \sigma\}$:
+\begin{equation*}
+\sigma(\mc{H}) := \inf\Biggl\{\sigma > 0: \inf_{\phi} \biggl[ \sup_{f_0 \in \mc{H}_{\sigma}} \Ebb_{f_0}[1 - \phi]\biggr] \leq \alpha\Biggr\}
+\end{equation*} 
+where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation under the regression function $f_0$.
+Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing critical radius over $\mc{H}$ is much smaller than the minimax estimation rate \citep{ingster2009}:
+\begin{equation}
+\label{eqn:sobolev_space_testing_critical_radius}
+\sigma^2(\mc{H}) \asymp M^{2d/(4s + d)}n^{-4s/(4s + d)}~~\textrm{for $1 \leq d < 4s$.}
+\end{equation}
+When $4s \geq d$ the functions in $H^s(\Xset)$ are very irregular---formally speaking $H^s(\Xset)$ does not \textcolor{red}{continuously} embed into $\Leb^4(\Xset)$ when $4s \geq d$---and the minimax testing rates in this regime are unknown.
+
+\textcolor{red}{(TODO)}
+\begin{itemize}
+	\item Is there a better citation than \citet{tsybakov2008_book} for the minimax rate of estimation~\eqref{eqn:minimax_estimation_rate}?
+Tsybakov considers the univariate case with fixed, equispaced design.
+	\item I copy pasted the testing part directly from our AISTATS paper, then made a few necessary changes. Is that bad?
+	\item I have stated the rate in~\eqref{eqn:sobolev_space_testing_critical_radius} over $\mc{H}$.
+In reality,~\citep{ingster2009} make some kind of boundary assumptions and I cannot remember exactly what.
+Is it ok to state it as is, or do I need to be more precise?
+\end{itemize}
+
+\subsection{Spectral series methods for Sobolev Spaces}
+Spectral series estimators are particularly designed for estimation in Sobolev spaces.
+Let $f_1,f_2,\ldots$ be the unit norm eigenfunctions of $\Delta_P$.
+The classical order-$K$ spectral series estimator $\wt{f}$ and test statistic $\wt{T}$ are
+\begin{equation*}
+\wt{f} = \sum_{k = 1}^{K} \dotp{Y}{f_k}_n f_k~~\textrm{and}~~\wt{T} = \|\wt{f}\|_P^2.
+\end{equation*}
+The estimator $\wt{f}$ and a test based on $\wt{T}$ are minimax rate-optimal over $H_0^s(\mc{X})$, a fact intimately related to the series representation of periodic Sobolev spaces, e.g.~\eqref{eqn:periodic_sobolev_space}.
+
+Unfortunately, in practice the estimator $\wt{f}$ is usually unworkable.
+Typically one does not have direct knowledge of the design distribution $P$, and even if $P$ is known, the eigenfunctions of $\Delta_P$ are rarely easy to compute.
+As an alternative, suppose instead we had access only to eigenfunctions $f_1',f_2',\ldots$ of the unweighted Laplace-Beltrami operator $\Delta$: then, letting $F_K = \mathrm{span}\{f_1,\ldots,f_K\}$, the least-squares estimator and test statistic
+\begin{equation*}
+\wt{f}_{\mathrm{LS}} := \argmin_{f \in F_K} \|Y - f\|_n^2,\textrm{and}~~\wt{T}_{\mathrm{LS}} = \|\wt{f}\|_{\nu}^2
+\end{equation*}
+are still rate-optimal over $H_0^s(P)$.
+However, this is not a totally satisfactory fix.
+For one thing, the least squares approach still requires that we know $\Delta$, and is not well suited for the case where $\mc{X}$ is a manifold, and $\Delta$ is the manifold Laplace-Beltrami operator.
+Additionally, diagonalizing even the unweighted Laplace-Beltrami operator $\Delta$ is quite difficult for all but a few special domains, such as $\mc{X} = [0,1]^d$.
+The general message is that in order to be minimax optimal, classical spectral series and least-squares estimators require an unrealistic amount of information on the design, and can be implemented only in selective situations.
+
+In contrast, Laplacian eigenmaps regression uses the eigenvectors of the neighborhood graph Laplacian $L_{n,\varepsilon}$ as features.
+The neighborhood graph Laplacian can be constructed (basically) without any knowledge of $P$ or $\mc{X}$, and Laplacian eigenmaps is thus a practicable method for regression.
+Moreover, the eigenvectors of $L_{n,\varepsilon}$ are approximations to the eigenfunctions $f_1,f_2,\ldots$ of $\Delta_P$---as we will discuss momentarily---so that $\wh{f}$ and $\wh{T}$ can in turn be viewed as approximations to the $\wt{f}$ and $\wt{T}$, respectively.
+As this view suggests, Laplacian eigenmaps naturally incurs some additional error compared to traditional spectral series estimators---this is the price we pay for using an approximation of $\Delta_P$.
+Fortunately, despite this extra error Laplacian eigenmaps retains the favorable optimality properties of its classical spectral series counterparts while freeing us from needing unrealistic knowledge of $\Delta_P$, as now summarize.
+
+\subsection{Our contributions}
+The following items summarize our major results.
+In all of them we assume~\ref{asmp:domain} and~\ref{asmp:density} hold, but not that any of $P$, $\mc{X}$, $\Delta_P$ or $\Delta$ are known.
+Put $r := s - 1$.
+\begin{itemize}
+	\item \textbf{Minimax optimal estimation.} For any $f_0 \in H^1(\Xset;M)$, with high probability $\norm{\wh{f} - f_0}_{n}^2 \lesssim M^{2d/(2 + d)}n^{-2/(2 + d)}$.
+	\item \textbf{Minimax optimal testing.}
+	A test constructed using $\wh{T}$ has non-trivial power whenever $f_0 \in H^1(\Xset;M)$ satisfies $\norm{f_0}_{P}^2 \gtrsim M^{2d/(2 + d)}n^{-4/(4 + d)}$ and $d < 4$.
+	\item \textbf{Higher-order smoothness conditions.} When the regression function $f_0 \in H_0^s(\Xset;M)$ and additionally the density $p \in C^{r}(\mc{X};M)$, with high probability the error $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim M^{2d/(2s + d)}n^{-2s/(2s + d)}$.
+Additionally, a test constructed using $\wh{T}$ has non-trivial power whenever $f_0 \in H_0^s(\Xset;M)$ satisfies $\norm{f_0}_{P}^2 \gtrsim M^{2d/(4s + d)}n^{-4s/(4s + d)}$ and $d < 4$.
+	\item \textbf{Manifold adaptivity.}
+	If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$ and $s \leq 2$, each of the aforementioned rates hold with $d$ replaced by $m$.
+\end{itemize}
+The above estimation results are stated with respect to in-sample mean squared error, meaning the loss is measured in the empirical squared-norm $\|\cdot\|_n^2$.\footnote{To state the obvious: this is very different than measuring the \emph{training error} $\|Y - \wh{f}\|_n^2$.} This is natural, as Laplacian eigenmaps is defined only at the observed design points $X_1,\ldots,X_n$.
+That being said, often one would like an upper bound on the prediction risk, or equivalently the out-of-sample loss $\|\cdot\|_P^2$.
+We therefore also introduce a simple, intuitive way of extending $\wh{f}$ to a function defined over all of $\mc{X}$, and show that this extension is minimax rate-optimal with respect to error measured in $L^2(P)$ norm.
+
+\subsection{Related Work}
+\begin{itemize}
+	\item \textcolor{red}{(TODO)} Summarize what is known on fixed graphs.
+	\item \textcolor{red}{Belkin, Slepcev} show that for a fixed $k$, the eigenvector $v_k$ converges to the eigenfunction $u_k$, in the sense that \textcolor{red}{(...)}.
+\textcolor{red}{Shi, Burago, Garcia Trillos, Calder, Cheng} build on this, giving finite sample bounds, rates of convergence, and making statements uniform over $1 \leq k \leq k_{\max}$.
+This justifies our previous discussion, in which we treat Laplacian eigenmaps as an approximation of spectral series methods.
+However, we emphasize that the minimax optimality of $\wh{f}$ or a test based on $\wh{T}$ does not follow straightforwardly from such results.
+Instead our analysis proceeds along different lines, which we detail in  Section~\ref{sec:analysis}.
+	\item There has been limited analysis of the estimator $\wh{f}$.
+\cite{zhou2011} consider a similar estimator, but in the semi-supervised setting where the number of unlabeled points grows to infinity.
+In this case the estimator reduces to a classical spectral series estimator, and there is no error incurred by approximating $\Delta_P$.
+\cite{lee2016} considers a similar estimator---using a different normalization of the Laplacian $L$---in both the supervised and semi-supervised setups, but get suboptimal rates for the supervised problem.
+To the best of knowledge, there has been no analysis of the test statistic $\wh{T}$ in the random design framework which we study.
+	\item \textcolor{red}{(TODO)}: Alternatively, \cite{trillos2020} \textcolor{blue}{(Green 2021)} use the graph Laplacian operator to induce a penalty over functions $f: \{\bf X\} \to \Reals$, and study a regularized estimator estimator, which minimizes the penalized loss
+	\begin{equation*}
+	\|Y - f\|_n^2 + \lambda \dotp{L_{n,\varepsilon}f}{f}_n.
+	\end{equation*}
+	\textcolor{blue}{(Green 2021)} show that the resulting estimator is (nearly)-minimax optimal, but only for $s = 1$ and $d \leq 4$.
+In contrast, the Laplacian eigenmaps estimator is optimal for all $s$ and $d$.
+\end{itemize}
+
+\section{Minimax Optimality of Laplacian Eigenmaps}
+\label{sec:minimax_optimal_laplacian_eigenmaps}
+
+In this section, we will show that the estimator $\wh{f}$, and a test using the statistic $\wh{T}$, achieve optimal estimation and goodness-of-fit testing rates over Sobolev classes.
+We will divide our statements based on whether $f_0$ belongs to the first order Sobolev class ($s = 1$) or a higher-order Sobolev class ($s > 1$), since the details of the two settings are somewhat different.
+Throughout this section, we will always assume~\ref{asmp:domain} and~\ref{asmp:density}.
+
+\subsection{First-order Sobolev classes}
+\label{sec:first_order_sobolev_classes}
+To begin, we handle the case where $f_0 \in H^1(\mc{X})$.
+
+\textbf{(1) Laplacian eigenmaps is an optimal estimator.} When the graph radius $\varepsilon$ and number of eigenvectors $K$ are chosen appropriately, the estimator $\wh{f}$ has (in-sample) optimal risk (up to constant factors) $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2/(2 + d)}$.
+\begin{enumerate}[label=(A\arabic*)]
+	\setcounter{enumi}{3}
+	\item 
+	\label{asmp:estimation_fo} 
+	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy the following inequalities:
+	\begin{equation}\\
+	\label{eqn:radius_estimation_fo} 
+	C_1 \biggl(\frac{\log n}{n}\biggr)^{1/d} \leq \varepsilon \leq \min\{i_0,K^{-1/d}\},
+	\end{equation}
+	and 
+	\begin{equation}
+	\label{eqn:eigenvector_estimation_fo} 
+	c_2 (M^2 n)^{2d/(2 + d)}\leq K \leq C_2 (M^2 n)^{2d/(2 + d)}.
+	\end{equation}
+\end{enumerate}
+\begin{theorem}
+	\label{thm:laplacian_eigenmaps_estimation_fo}
+	Suppose the general assumptions~\ref{asmp:domain}-\ref{asmp:sobolev_radius} are satisfied.
+Then for any $f_0 \in H^1(P,M)$, if Laplacian eigenmaps is run with parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:estimation_fo}, with probability at least $1 - \delta - \textcolor{red}{(???)}$ it follows that
+	\begin{equation}
+	\label{eqn:laplacian_eigenmaps_estimation_fo}
+	\|\wh{f} - f_0\|_n^2 \leq \frac{C}{\delta}M^{2d/(2 + d)}n^{-2/(2 + d)}
+	\end{equation}
+\end{theorem}
+Some remarks:
+\begin{itemize}
+	\item \textbf{(1.1) How to tune hyper-parameters.} We note that the ranges~\eqref{eqn:radius_estimation_fo} and~\eqref{eqn:eigenvector_estimation_fo} depend on unknown quantities such as $f_0$ and $P$.
+In practice, one typically tunes hyper-parameters by sample-splitting or cross-validation.
+However, because the estimator $\wh{f}$ is defined only in-sample, such approaches cannot be straightforwardly applied to select the graph radius $\varepsilon$, or number of eigenvectors $K$.
+We discuss out-of-sample extensions of $\wh{f}$ later, in Section~\ref{sec:out_of_sample}.
+	\item \textbf{(1.2) The dependence on failure probability can be improved.} The upper bound on in-sample risk given by Equation~\eqref{eqn:laplacian_eigenmaps_estimation_fo} depends on the pre-factor $C/\delta$, and holds with probability $1 - \delta - \textcolor{red}{(?)}$.
+When $f_0 \in C^1(\mc{X};M)$, we can improve these dependencies, changing the pre-factor in~\eqref{eqn:laplacian_eigenmaps_estimation_fo} to $C(1 + 1/\delta)$, and showing that the bound holds with probability at least $1 - C\delta^2/n$.
+\end{itemize}
+
+\textbf{(2) Laplacian eigenmaps is an optimal test.} Consider the test $\varphi = \1\{\wh{T} \geq t_{\delta}\}$, where $t_{\delta}$ is the threshold
+\begin{equation*}
+t_{\delta} := \frac{K}{n} + \frac{2K}{\sqrt{n}}.
+\end{equation*}
+The choice of $t_{\delta}$ guarantees that $\varphi$ is a level-$\delta$ test.
+Under appropriate choices of $\varepsilon$ and $K$, the test $\varphi$ has non-negligible power (up to constant factors) against alternatives separated the null by at least $\|f_0\|_{P}^2 \gtrsim M^{2d/(4 + d)}n^{-4/(4 + d)}$.
+
+\begin{theorem}
+	\label{thm:laplacian_eigenmaps_testing_fo}
+	\textcolor{red}{(TODO)}
+\end{theorem}
+Some remarks:
+\begin{itemize}
+	\item \textbf{(2.1) When $d \geq 4$, the testing problem is different.} When $d \geq 4$, it is not the case that $\Leb^4(\mc{X})$ continuously embeds into $H^1(\mc{X})$, and testing rates over $H^1(\mc{X})$ are not known.
+On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then simply taking $T = \|Y\|_n^2$ achieves the optimal rates.
+	\item \textbf{(2.2) Both of the remarks after Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} also apply to testing.}
+\end{itemize}
+
+\subsection{Higher-order Sobolev classes}
+\label{sec:higher_order_sobolev_classes}
+In this section, we now assume that the regression function displays some higher-order regularity, $f_0 \in H_0^s(\mc{X})$.
+We point out that on a graph there is no natural analogue to higher-order derivatives.
+For this reason it is not at all obvious that graph-based methods \emph{can} adapt to higher-order smoothness.
+Remarkably, we show that this is in fact the case.
+Under appropriate choices of $\varepsilon$ and $K$, the estimator $\wh{f}$ has (in-sample) optimal error (up to constant factors) $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2s/(2s + d)}$.
+\begin{theorem}
+	\label{thm:laplacian_eigenmaps_estimation_ho}
+	Under appropriate conditions, with probability at least $1 - \delta$,
+	\begin{equation}
+	\label{eqn:laplacian_eigenmaps_estimation_ho}
+	\|\wh{f} - f_0\|_n^2 \lesssim n^{-2s/(2s + d)}
+	\end{equation}
+\end{theorem}
+Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that in the flat Euclidean setting Laplacian eigenmaps is a rate-optimal estimator over Sobolev classes, for all scalings of $s$ and $d$.
+Some other remarks:
+\begin{itemize}
+	\item \textbf{(1.1) Optimal rates, no RKHS.} It is worth pointing out that we do not require that the regularity of the Sobolev space satisfy $2s > d$, a condition often seen in the literature.
+In the ``sub-critical'' regime $2s < d$, the Sobolev space $H^s(\mc{X})$ is quite irregular.
+For instance, it is not a Reproducing Kernel Hilbert Space, nor does it embed into any H\"{o}lder space.
+As a result, one has to be a bit careful.
+Certain versions of the nonparametric regression problem are ill-posed---for instance when loss is measured in $\Leb^{\infty}$ norm, or when the design points $X$ are assumed to be fixed.
+It also takes certain estimators ``off the table'', most notably RKHS-based methods such as smoothing splines.
+However, at these scales spectral series estimators still obtain the optimal rates for random design regression measured in $\Leb^2(P)$ norm.
+Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} shows that the same is true with respect to Laplacian eigenmaps, with error measured in $\Leb^2(P_n)$ norm.
+	\item \textbf{(1.2) The zero-trace condition.} The zero-trace condition can likely be weakened.
+However, requiring some boundary condition on $f_0$ and its first through $(k -1 )$th-order derivatives is unavoidable.
+\textcolor{red}{Explain why: graphs are implicitly boundaryless objects, and the eigenvectors of $L$ converge to those eigenvectors of $\Delta_P$ which satisfy boundary conditions.
+Repeat that this is a standard assumption made when analyzing spectral series estimators}.
+	\item \textbf{(1.3) The bounded norm and smooth density conditions.} We require that $f_0$ be bounded in the Sobolev norm $\|f_0\|_{H^s(\mc{X})}$---as opposed to the semi-norm $|f_0|_{H^s(\mc{X})}$--- and that $p \in C^{s - 1}(\mc{X})$.
+Both requirements are related and fundamental, as we shall now explain.
+As previously mentioned, $L$ converges to a weighted-version of the Laplace operator $\Delta_P$ (also known as a Fokker-Planck) operator.
+The operator $\Delta_P$ includes both first and second-order derivatives,
+	\begin{equation}
+	\label{eqn:fokker_planck_1}
+	\Delta_Pf= \frac{-1}{p} \mathrm{div}(p^2 \nabla f) = \Delta f - 2\frac{\nabla p^{\top} \nabla f}{p},
+	\end{equation}
+	and therefore a function $f \in H_0^s(\mc{X})$ has finite energy $\|\Delta_P f\|_{P}^2$ only if the first-order partial derivatives of $f$ and $p$ are bounded.
+\textcolor{red}{This discussion---and in particular the last sentence---are obviously not quite correct.
+This is in part because you realize now that the requirements are not in fact fundamental.
+Instead, the requirement should merely be that $\|\Delta_P^kf\|_{\Leb^2(P)}^2$ is bounded---here $k = s/2$ and we assume $s$ is even---and the upper bounds should be in terms of this semi-norm.}.
+	\item \textbf{(1.4) Both of the remarks after Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} continue to apply.}
+\end{itemize}
+
+\textbf{(1.4) Rates for testing with higher-order smoothness assumptions.} For testing, we can get the optimal rate $n^{-4s/(4s + d)}$ when $d \leq 4$.
+When $4 < d \leq 4s$, we get the rate~\textcolor{red}{(...)} Recall that the rate $n^{-4s/(4s + d)}$ is optimal whenever $d \leq 4s$.
+Therefore there are scalings of $s$ and $d$ for which we do not get the optimal rate.
+We do not know whether this due to looseness in our proof techniques, or a problem with the test itself.
+
+\subsection{Analysis}
+\label{sec:analysis}
+
+\textbf{(1) Bias and variance terms, for estimation and testing}.
+We analyze the estimation error of $\wh{f}$ by first condition on the design points $X$ and computing \emph{design-dependent} bias and variance terms.
+For estimation, we have that with high probability, conditional on the design $X$,
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 \leq C\frac{\dotp{L^s f_0}{f_0}_n}{\lambda_{\kappa}^s} + \frac{K}{n}
+\end{equation*}
+For testing, we have~\textcolor{red}{(...)}.
+
+The semi-norm $\dotp{L^s f_0}{f_0}_n$ and eigenvalue $\lambda_{K}$ are random variables that depend the design $X$. It remains to establish suitable upper and lower bounds on these quantities.
+
+\textbf{(2) Upper bound on the semi-norm. } Proposition~\ref{prop:graph_seminorm} gives a sufficient upper bound on $\dotp{L^s f_0}{f_0}_n$. Take $\beta := s - 1$. 
+\begin{proposition}
+	\label{prop:graph_seminorm} 
+	Suppose $f \in H_0^s(\mc{X})$, that $p \in C^{\beta}(\mc{X})$, and that $r \gtrsim n^{-1/d}$. Then with probability at least $1 - \delta$, it holds that 
+	\begin{equation}
+	\label{eqn:graph_seminorm_1}
+	\dotp{L^s f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^s} \|p\|_{C^{\beta}} + \textcolor{red}{(?)}
+	\end{equation}
+	Therefore, if additionally $r \gtrsim n^{-1/(2\beta + d)}$, then 
+	\begin{equation}
+	\label{eqn:graph_seminorm_2}
+	\dotp{L^s f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^s} \|p\|_{C^{\beta}}
+	\end{equation}
+	When $s = 1$, the results hold for $f \in H^1(\mc{X})$ and any density $p$ that satisfies Assumption~\ref{asmp:density}.
+\end{proposition}
+The proof of Proposition~\ref{prop:graph_seminorm} involves two steps.
+In the first step, we show that with probability at least $1 - \delta$, the graph semi-norm $\dotp{L^s f}{f}_n$ is upper bounded by a non-local continuum energy $E_{\varepsilon}^{(s)}(f)$.
+This non-local continuum energy is defined as the following semi-norm,
+\begin{equation*}
+E_{\varepsilon}^{(s)}(f) := \dotp{L_{P,\varepsilon}^sf}{f}_{P},
+\end{equation*}
+where $L_{P,\varepsilon}$ is a non-local approximation to $\Delta_P$, 
+\begin{equation*}
+L_{P,\varepsilon}f(x) := \int \bigl(f(z) - f(x)\bigr) \eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dP(x).
+\end{equation*}
+In the second step, we show that $E_{\varepsilon}^{(s)}(f)$ is in turn upper bounded by a constant times $D_s(f) := \|f\|_{H^s(P)}^2$.
+Together, these steps establish Proposition~\ref{prop:graph_seminorm}.
+
+When $s = 1$, these results are somewhat standard~\textcolor{red}{(references)}.
+On the other hand, much less work has been done in the $s > 1$ case, and the analysis is more complicated.
+For one thing, we note that when $s > 1$ the graph energy $\dotp{L^s f}{f}_n$ is itself a biased estimate of the non-local energy $E_{\varepsilon}^{(s)}(f)$.
+This bias term is the second term on the right hand side of~\eqref{eqn:graph_seminorm_1}, and is neglible when $r \gtrsim n^{-1/(2r + d)}$.
+On the other hand, in order to give an estimate of $E_{\varepsilon}^{(s)}(f)$ in terms of the Sobolev norm $D_s(f)$, we first show that $L_{P,\varepsilon}^jf(x) \approx \Delta_P^jf(x)$, for $j = s/2$ when $s$ is odd and $j = (s - 1)/2$ when $s$ is even.
+In words, we show that a $j$th-order iterated difference operator approximates a $2j$-th order differential operator.
+It is in order to establish this (somewhat remarkable) fact that we require on the one hand $p \in C^{r}(\mc{X};M)$, and on the other hand that $f$ be zero-trace.
+
+\textbf{(3) Lower bound on the eigenvalue.} On the other hand, recent work \textcolor{red}{(Burago14, Shi16, Garcia Trillos 18)} has analyzed the convergence of $\lambda_{k}$ towards $\lambda_{k}(\Delta_P)$.
+They provide explicit bounds on the relative error $|\lambda_{k} - \lambda_{k}(\Delta_P)|/\lambda_{k}(\Delta_P)$, showing that the relative error is small for sufficiently large $n$, small $\varepsilon$, and any $1 \leq k \leq n$ that is not too large relative to $n$.
+These results are actually stronger than are necessary to establish Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_estimation_ho}---in order to get rate-optimality, we need only show that $\lambda_{K}/\lambda_K(P)$ is $\Omega_P(1)$---but unfortunately they all assume $P$ is supported on a manifold without boundary.
+We will instead use the results of \textcolor{red}{(Green 2021)}, whose assumptions match our own, and who give a weaker bound on $\lambda_k/\lambda_k(\Delta_P)$ that will nevertheless suffice for our purposes.
+
+\begin{proposition}
+	\textcolor{red}{(TODO)}
+	With probability at least $1 - \textcolor{red}{(?)}$,
+	\begin{equation*}
+	\lambda_k \geq c \cdot \min\Bigl\{\lambda_k(\Delta_P), \frac{1}{r^{2}} \Bigr\}
+	\end{equation*}
+\end{proposition}
+By Weyl's Law, $c k^{2/d} \leq \lambda_{k}(\Delta_P) \leq Ck^{2/d}$, and as a result $\lambda_{K} \geq C\lambda_{K}(\Delta_P)$ as long as $r \geq \kappa^{-1/d}$.
+
+\textbf{(1.2) Traps we avoid.} It is tempting to use recent results regarding spectral convergence to analyze the estimator $\wh{f}$ and test $\wh{T}$.
+This suffers from an accumulation of error problem.
+
+It is also key that we directly analyze the semi-norm $\dotp{L^s f}{f}_n$, rather than invoking the pointwise convergence of $L^{s}f \to \Delta_P^{s}f$ to obtain a bound on $\dotp{L^s f}{f}_n$.
+Recall that we assume $f$ has bounded derivatives up to (and including) order $s$.
+This suffices to gain control of the semi-norm $\dotp{L^s f}{f}_n$, because $\dotp{L^s f}{f}_n$ approximates the $\Leb^2$ norm of an order $s$ differential operator, and so we have exactly as many derivatives as we need.
+On the other hand, $L^s$ approximates an order $2s$ differential operator, and $\Delta_P^sf$ is only a well-defined limiting object if we assume $f$ has order $2s$ derivatives.
+
+\textcolor{red}{(TODO)}
+\begin{itemize}
+	\item Possibly turn the first part of the discussion after Proposition~\ref{prop:graph_seminorm} into a small proof.
+\end{itemize}
+
+\section{Manifold Adaptivity}
+\label{sec:manifold_adaptivity}
+
+The optimal rates for nonparametric regression suffer from a bona fide curse of dimensionality, and are thus difficult to reconcile with the practical success of many methods for nonparametric regression.
+One explanation that resolves this seeming paradox is the \emph{manifold hypothesis}, where the design points $X$ are assumed to lie on or near a manifold $\mc{X}$ of intrinsic dimension $m \ll d$.
+Under the manifold hypothesis nonparametric regression is demonstrably easier, and in particular it is known \citep{bickel2007,ariascastro2018} that in this setting the optimal rates over $H^s(\Xset)$ scale like $n^{-2s/(2s + m)}$ (for estimation) and $n^{-4s/(4s + m)}$ (for testing).
+
+On the other hand, a theory has been developed~\citep{belkin03,belkin05,niyogi2013} establishing the the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+In this section, we contribute to this line of work by showing that under the manifold hypothesis, Laplacian eigenmaps achieve the sharper minimax estimation and testing rates.
+First, we give some relevant background on Riemmanian manifolds, and Sobolev classes defined upon them.
+
+\subsection{Riemmanian manifolds}
+We assume $\mc{X}$ is a closed, connected, smooth manifold without boundary of dimension $m$ isometrically embedded in $\Rd$.
+We give to $\mc{X}$ the Riemannian structure induced by the ambient space $\Rd$, and denote the resulting volume form by $\mu$.
+
+\textcolor{red}{(TODO)} Other objects to mention, although probably some of them need only be described in the appendix.
+\begin{itemize}
+	\item Exponential map, injectivity radius, reach, Laplace-Beltrami operator.
+\end{itemize}
+
+\subsection{Error rates under the manifold hypothesis}
+
+\section{Out-of-sample error}
+\label{sec:out_of_sample}
+Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity} show that Laplacian eigenmaps is an optimal method for nonparametric regression.
+As explained previously, we have measured estimation error in $\|\cdot\|_n$ norm---in words, the average squared loss incurred over the $n$ observed design points--- whereas error measured in $L^2(P)$ norm is the more typical metric in the random design setup.
+
+Of course, the Laplacian eigenmaps estimator is only defined at the observed design points $X_1,\ldots,X_n$, and to measure its error in $L^2(P)$ norm we must first give a means of extending it to the rest of $\Xset$.
+We propose a simple method, essentially kernel smoothing, to do the job, which can be applied to any estimator defined at the design points, including Laplacian eigenmaps.
+We show that the smoothed version of $\wh{f}$ has optimal $L^2(P)$ error.
+
+\subsection{Extension using Kernel Smoothing}
+
+\subsection{Out-of-sample error of Kernel Smoothed Laplacian Eigenmaps}
+
+\begin{theorem}
+	\label{thm:kernel_smoothing}
+	\textcolor{red}{(TODO)}
+\end{theorem}
+
+Some remarks:
+\begin{itemize}
+	\item \textbf{Tuning parameters by sample-splitting.}
+	\item \textbf{Other extensions.} Comment that there does not exist any canonical way to extend Laplacian based estimators such as $\wh{\theta}$ to an ambient domain.
+Point out that there exist various proposals to extend eigenvectors in the literature, such as $1$-nearest-neighbor regression (piecewise constant extrapolation over Voronoi cells), convolution with a specialized kernel, or Nystr\"{o}m extension.
+Explain that each of these is designed for a highly specialized purpose, such as suitability for mathematical analysis or computational efficiency, and in any case are suggested only for eigenvectors rather than a regression estimate.
+State explicitly that we choose kernel smoothing for its simplicity, ubiquity, and strong theoretical properties under our assumptions, but that considerations of extensions for specifically tailored to Laplacian based estimators could be of interest.
+\end{itemize}
+
+\subsection{Why not just do kernel smoothing?}
+On the other hand, the allusion to kernel smoothing raises a very natural question: why should we prefer to first compute $\wh{\theta}$ and then smooth it out, as opposed to simply performing kernel smoothing over the original responses $Y$?
+
+We now answer this question, by \textcolor{red}{(...)}.
+
+\section{Experiments}
+\label{sec:experiments}
+
+\textcolor{red}{(TODO)}: Come up with a set of experiments that might be interesting to run. Run them by Ryan and Siva, so that you can prioritize.
+
+\section{Discussion}
+\label{sec:discussion}
+
+\textbf{(1) Future items:}
+\begin{itemize}
+	\item \textbf{()}
+\end{itemize}
+
 \bibliographystyle{plainnat}
-\bibliography{../../graph_testing_bibliography}
+\bibliography{../../graph_regression_bibliography}
 

### 2021-02-23 manuscripts/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
Detailed pass over the main AISTATS paper

--- 2020-09-28 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing.tex
@@
-\documentclass{article}
-\usepackage{amsmath}
-\usepackage{amsfonts, amsthm, amssymb}
+\documentclass[twoside]{article}
+\usepackage[accepted]{aistats2021}
+\usepackage{amsmath, amsfonts, amsthm, amssymb}
 \usepackage{graphicx}
-\usepackage[colorlinks]{hyperref}
+\usepackage{hyperref}
+\hypersetup{
+	colorlinks=true,
+	linkcolor=blue,
+	citecolor=blue
+}
 \usepackage[parfill]{parskip}
@@
 \usepackage[shortlabels]{enumitem}
-\usepackage{fullpage}
 \usepackage{mathtools}
 \usepackage{tikz}
+\usepackage{verbatim}
+\usepackage{microtype}
+\usepackage{bm}
 
@@
 \DeclareMathOperator*{\argmax}{argmax}
+\DeclareMathOperator*{\esssup}{ess\,sup}
 
@@
 \newtheorem{definition}{Definition}[section]
-
 \theoremstyle{remark}
@@
 \begin{document}
-\title{Minimax optimal Laplacian smoothing}
-\author{Alden Green}
-\date{\today}
-\maketitle
 
-\textbf{Disclaimer:} This largely follows the same setup as in the draft I have written which combines Laplacian smoothing and Laplacian eigenmaps results.
-That draft --- document name \textit{graph\_regression2} --- should for the moment be considered obsolete.
+\runningtitle{Minimax Optimal Laplacian Smoothing}
 
-\section{Introduction}
+\twocolumn[
 
-\textbf{(1)} Laplacian smoothing is a graph-based approach to nonparametric regression.
+	\aistatstitle{Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs}
 
+	\aistatsauthor{Alden Green \And Sivaraman Balakrishnan \And Ryan J. Tibshirani}
+
+	\aistatsaddress{Carnegie Mellon University \And Carnegie Mellon University \And Carnegie Mellon University}]
+
+\begin{abstract}
+	In this paper we study the statistical properties of Laplacian smoothing, a graph-based approach to nonparametric regression.
+Under standard regularity conditions, we establish upper bounds on the error of the Laplacian smoothing estimator \smash{$\wh{f}$}, and a goodness-of-fit test based on \smash{$\wh{f}$}.
+These upper bounds match the minimax optimal estimation and testing rates of convergence over the first-order Sobolev class $H^1(\Xset)$, for $\Xset \subseteq \Rd$ and $1 \leq d < 4$; in the estimation problem, for $d = 4$, they are optimal modulo a $\log n$ factor.
+Additionally, we prove that Laplacian smoothing is manifold-adaptive: if $\Xset \subseteq \R^d$ is an $m$-dimensional manifold with $m < d$, then the error rate of Laplacian smoothing (in either estimation or testing) depends only on $m$, in the same way it would if $\Xset$ were a full-dimensional set in $\Reals^m$.
+\end{abstract}
+
+\section{INTRODUCTION}
+
+We adopt the standard setup in nonparametric regression, where we observe samples $(X_1,Y_1),\ldots,(X_n,Y_n)$ that are i.i.d.\ draws from the model 
+\begin{equation}
+\label{eqn:signal_plus_noise_model}
+Y_i = f_0(X_i) + \varepsilon_i, \quad \epsilon_i \sim N(0,1), 
+\end{equation}
+where $\varepsilon_i$ is independent of $X_i$.
+Our goal is to perform statistical inference on the unknown regression function $f_0$, by which we mean either \emph{estimating} $f_0$ or more simply \emph{testing} whether $f_0 = 0$, i.e., whether there is any signal present.
+
+Laplacian smoothing \citep{smola2003} is a penalized least squares estimator, defined over a graph.
+Let $G = (V,W)$ be a weighted undirected graph with vertices $V=\{1,\ldots,n\}$, associated with $\{X_1,\ldots,X_n\}$, the Laplacian smoothing estimator \smash{$\wh{f}$} is given by
+\begin{equation}
+\label{eqn:laplacian_smoothing}
+\wh{f} =  \argmin_{f \in \Reals^n} \; \sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^\top \Lap f. 
+\end{equation}
+Here $\Lap$ is the graph Laplacian matrix (to be defined in Section~\ref{sec:problem_setup_and_background}), $G$ is typically a geometric graph (such as a $k$-nearest-neighbor or neighborhood graph), $\rho \geq 0$ is a tuning parameter, and the penalty
+\begin{equation*}
+f^\top \Lap f = \frac{1}{2} \sum_{i,j = 1}^{n} W_{ij}(f_i - f_j)^2
+\end{equation*}
+encourages \smash{$\wh{f}_i \approx \wh{f}_j$} when $X_i \approx X_j$. Assuming \eqref{eqn:laplacian_smoothing} is a reasonable estimator of $f_0$, the statistic
+\begin{equation}
+\label{eqn:laplacian_smoothing_test}
+\wh{T} = \frac{1}{n} \| \wh{f} \|_2^2 
+\end{equation}
+would in turn be a reasonable way to test $f_0 = 0$.
+
+Of course there are many methods for nonparametric regression (see, e.g., \citet{gyorfi2006,wasserman2006,tsybakov2008_book}), but Laplacian smoothing has its own set of advantages. For instance:
 \begin{itemize}
-	\item In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Rd$, and
+	\item \emph{Computational ease.} Laplacian smoothing is fast, easy, and stable to compute.
+The estimate \smash{$\wh{f}$} can be computed by solving a symmetric diagonally dominant linear system.
+There are by now various nearly-linear-time solvers for this problem (see e.g., the seminal papers of \citet{spielman2011,spielman2013,spielman2014}, or the overview by \citet{vishnoi2012} and references therein).
+	\item \emph{Generality.} Laplacian smoothing is well-defined whenever one can associate a graph with observed responses.
+This generality lends itself to many different data modalities, e.g., text and image classification, as in \citet{kondor2002,belkin03a,belkin2006}.
+	\item \emph{Weak supervision.} Although we study Laplacian smoothing in the supervised problem setting \eqref{eqn:signal_plus_noise_model}, the method can be adapted to the semi-supervised or unsupervised settings, as in \citet{zhu2003semisupervised,zhou2005learning,nadler09}.
+%,slepcev17,calder2019b,dunlop2020}.
+\end{itemize}
+
+For these reasons, a body of work has emerged that analyzes the statistical properties of Laplacian smoothing, and graph-based methods more generally.
+Roughly speaking, this work can be divided into two categories, based on the perspective they adopt.
+\begin{itemize}
+	\item \emph{Fixed design perspective.} Here one treats the design points $X_1,\ldots,X_n$ and the graph $G$ as fixed, and carries out inference on $f_0(X_i)$, $i=1,\ldots,n$.
+In this problem setting, tight upper bounds have been derived on the error of various graph-based methods (e.g., \citet{wang2016,hutter2016,sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}) and tests (e.g., \citet{sharpnack2010identifying,sharpnack2013b,sharpnack2013,sharpnack2015}), which certify that such procedures are \emph{optimal} over ``function'' classes (in parentheses because these classes really model the $n$-dimensional vector of evaluations).
+The upside of this work is its generality: in this setting $G$ need not be a geometric graph, but in principle it could be any graph over $V=\{1,\ldots,n\}$.
+The downside is that, in the context of nonparametric regression, it is as not natural to think of the evaluations of $f_0$ as exhibiting smoothness over some fixed pre-defined graph $G$, and more natural to speak of the smoothness of the function $f_0$ itself.
+	\item \emph{Random design perspective.} Here one treats the design points $X_1,\ldots,X_n$ as independent samples from some distribution $P$ supported on a domain $\Xset \subseteq \Rd$.
+Inference is drawn on the regression function $f_0: \Xset \to \Reals$, which is typically assumed to be smooth in some \emph{continuum} sense, e.g., it possesses a first derivative bounded in $L^{\infty}$ (H\"{o}lder) or $\Leb^2$ (Sobolev) norm.
+To conduct graph-based inference, the user first builds a neighborhood graph over the random design points---so that $W_{ij}$ is large when $X_i$ and $X_j$ are close in (say) Euclidean distance---and then computes e.g., \eqref{eqn:laplacian_smoothing} or \eqref{eqn:laplacian_smoothing_test}.
+In this context, various graph-based procedures have been shown to be \emph{consistent}: as $n \to \infty$, they converge to a continuum limit (see \citet{belkin07,vonluxburg2008,trillos2018} among others).
+However, until recently such statements were not accompanied by error rates, and even so, such error rates as have been proved \citep{lee2016,trillos2020} are not optimal over continuum function spaces, such as H\"{o}lder or Sobolev classes.
+\end{itemize}
+The random design perspective bears a more natural connection with nonparametric regression (the focus in this paper), as it allows us to formulate smoothness based on $f_0$ itself (how it behaves as a continuum function, and not just its evaluations at the design points).
+In this paper, we will adopt the random design perspective, and seek to answer the following question:
+\begin{quote}{When $f_0$ is assumed to smooth in a continuum sense, does Laplacian smoothing achieve optimal performance for estimation and goodness-of-fit testing?} 
+\end{quote}
+This is no small question---arguably, it is \emph{the} central question of nonparametric regression---and without an answer one cannot fully compare the statistical properties of Laplacian smoothing to alternative methods.
+It also seems difficult to answer: as we discuss next, there is a fundamental gap between the \emph{discrete} smoothness imposed by the penalty $f^\top \Lap f$ in problem \eqref{eqn:laplacian_smoothing} and the \emph{continuum} smoothness assumed on $f_0$, and in order to obtain sharp upper bounds we will need to bridge this gap in a suitable sense.
+
+\section{SUMMARY OF RESULTS}
+
+\paragraph{Advantages of the Discrete Approach.}
+
+In light of this last point, it is worth asking whether there is any \emph{statistical} advantage to solving a discrete problem such as \eqref{eqn:laplacian_smoothing} (setting aside computational considerations for the moment).
+After all, we could have instead solved the following variational problem:
 \begin{equation}
-	\label{eqn:random_design_regression}
-	Y_i = f_0(X_i) + \varepsilon_i
+\label{eqn:thin_plate_spline}
+\wt{f} = \argmin_{f : \Xset \to \Reals} \; \sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \hspace{-2pt} \int_{\Xset} \|\nabla f(x)\|_2^2 \,dx,
 \end{equation}
-	with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise.
-Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present.
-	\item In graph-based nonparametric regression, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$.
-The neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$.
-The $n \times n$ weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $\mathbf{W}_{ij}$ are given by
+where the optimization is performed over all functions for which the criterion is well-defined and finite (continuous functions $f$ that have a weak derivative $\nabla f$ in $L^2(\Xset)$).
+Analogously, for testing, we could use:
+\begin{equation}
+\label{eqn:thin_plate_spline_test}
+\wt{T} = \| \wt{f} \|_n^2 = \frac{1}{n} \sum_{i=1}^n \wt{f}(X_i)^2.
+\end{equation}
+The penalty term in \eqref{eqn:thin_plate_spline} leverages the assumption that $f_0$ has a smooth derivative in a seemingly natural way.
+Indeed, the estimator \smash{$\wt{f}$} and statistic \smash{$\wt{T}$} are well-known: for $d = 1$, \smash{$\wt{f}$} is the familiar \emph{smoothing spline}, and for $d>1$, it is type of a \emph{thin-plate spline}.
+The statistical properties of smoothing and thin-plate splines are well-understood \citep{vandergeer2000,liu2019}.
+As we discuss later, the Laplacian smoothing problem \eqref{eqn:laplacian_smoothing} can be viewed as a discrete and noisy approximation to \eqref{eqn:thin_plate_spline}.
+At first blush, this suggests that Laplacian smoothing should at best inherit the statistical properties of \eqref{eqn:thin_plate_spline}, and at worst may have meaningfully larger error.
+
+However, as we shall see the actual story is quite different: remarkably, Laplacian smoothing enjoys optimality properties even in settings where the thin-plate spline estimator \eqref{eqn:thin_plate_spline} is not well-posed (to be explained shortly); Tables~\ref{tbl:estimation_rates} and~\ref{tbl:testing_rates} summarize.
+As we establish in Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing_manifold}, when computed over an appropriately formed neighborhood graph, Laplacian smoothing estimators and tests are minimax optimal over first-order \emph{continuum} Sobolev balls.
+This holds true either when $\Xset \subseteq \Rd$ is a full-dimensional domain and $d = 1,2,$ or $3$, or when $\Xset$ is a manifold embedded in $\Rd$ of intrinsic dimension $m = 1,2,$ or $3$.
+Additionally, the estimator \smash{$\wh{f}$} is nearly minimax optimal (to within a \smash{$(\log n)^{1/3}$} factor) when $d=4$ (or $m=4$ in the manifold case).
+
+By contrast, smoothing splines are optimal only when $d=1$.
+When $d>1$, thin-plate spline estimator \eqref{eqn:thin_plate_spline} is not even well-posed, in the following sense: for any $(X_1,Y_1),\ldots,(X_n,Y_n)$ and any $\delta>0$, there exists (e.g., \citet{green93} give a construction using ``bump'' functions) a differentiable function $f$ such that $f(X_i) = Y_i$, $i=1,\ldots,n$, and
+\begin{equation*}
+\int_{\Xset} \|\nabla f(x)\|_2^2 \leq \delta.
+\end{equation*}
+i.e., $f$ achieves perfect (zero) data loss and arbitrarily small penalty in the problem \eqref{eqn:thin_plate_spline}.
+This will clearly not lead to a consistent estimator of $f_0$ across the design points (as it always yields $Y_i$ at each $X_i$).
+In this light, our results for $d>1$ favorably distinguish Laplacian smoothing from its natural variational analog.
+
+\begin{table}
+	\begin{center}
+		\begin{tabular}{p{.1\textwidth} | p{.14\textwidth} p{.12\textwidth} }
+			Dimension & Laplacian smoothing \eqref{eqn:laplacian_smoothing} & Thin-plate splines \eqref{eqn:thin_plate_spline} \\
+			\hline
+			$d = 1$ & $\bm{n^{-2/3}}$ & \textcolor{red}{$\bm{n^{-2/3}}$} \\
+			$d = 2,3$ & $\bm{n^{-2/(2 + d)}}$ & \textcolor{red}{$1$} \\
+			$d = 4$ & $\bm{n^{-1/3}} (\log n)^{1/3}$ & \textcolor{red}{$1$} \\
+			$d \geq 5$  & $(\log n/n)^{4/(3d)}$ &\textcolor{red}{$1$} \\
+		\end{tabular}
+	\end{center}
+	\caption{Summary of estimation rates over first-order Sobolev balls.
+Black font marks new results from this paper, red font marks previously-known results; bold font marks minimax optimal rates.
+Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius of the Sobolev ball is also optimal.
+We use ``$1$'' to indicate inconsistency (error not converging to 0).
+Furthermore, when $\Xset$ is an $m$-dimensional manifold embedded in $\Rd$, all Laplacian smoothing results hold with $d$ replaced by $m$, without any change to the method itself.}
+	\label{tbl:estimation_rates}
+\end{table}
+
+\begin{table}
+	\begin{center}
+		\begin{tabular}{p{.1\textwidth} | p{.14\textwidth} p{.12\textwidth} }
+			Dimension & Laplacian smoothing \eqref{eqn:laplacian_smoothing_test} & Thin-plate splines \eqref{eqn:thin_plate_spline_test} \\
+			\hline
+			$d = 1$ & $\bm{n^{-4/5}}$ & \textcolor{red}{$\bm{n^{-4/5}}$} \\
+			$d = 2,3$ & $\bm{n^{-4/(4 + d)}}$ & \textcolor{red}{$n^{-1/2}$} \\
+			$d \geq 4$ & $\bm{n^{-1/2}}$ & \textcolor{red}{$\bm{n^{-1/2}}$}
+		\end{tabular}
+	\end{center}
+	\caption{Summary of testing rates over first-order Sobolev balls; black, red, and bold fonts are used as in Table~\ref{tbl:estimation_rates}.
+Rates for $d \geq 4$ assume that $f_0 \in \Leb^4(\Xset,M)$.
+When $\Xset$ is an $m$-dimensional manifold embedded in $\Rd$, all rates hold with $d$ replaced by $m$.}
+	\label{tbl:testing_rates}
+\end{table}
+%\footnotetext{Assuming $f_0 \in \Leb^4(\Xset,M)$.}
+
+\paragraph{Future Directions.}
+
+To be clear, there is still much left to be investigated.
+For one, the Laplacian smoothing estimator \smash{$\wh{f}$} is only defined at $X_1,\ldots,X_n$.
+In this work we study its in-sample mean squared error
+\begin{equation}
+\label{eqn:empirical_norm_error}
+\bigl\|\wh{f} - f_0\bigr\|_n^2 := \frac{1}{n}\sum_{i = 1}^{n} \Bigl(\wh{f}_i - f_0(X_i)\Bigr)^2.
+\end{equation}
+In Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we discuss how to extend \smash{$\wh{f}$} to a function over all $\Xset$, in such a way that the out-of-sample mean squared error \smash{$\|\wh{f} - f_0\|_{\Leb^2(\Xset)}^2$} should remain small, but leave a formal analysis to future work.
+
+In a different direction, problem \eqref{eqn:thin_plate_spline} is only a special, first-order case of thin-plate splines. In general, the $k$th order thin-plate spline estimator is defined as
+\begin{equation*}
+\wt{f} = \argmin_{f : \Xset \to \Rd} \; \sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \hspace{-2pt} \sum_{|\alpha| = k} \int_{\Xset} \bigl(D^\alpha f(x)\bigr)^2 \,dx,
+\end{equation*}
+where for each multi-index $\alpha=(\alpha_1,\ldots,\alpha_d)$ we write $D^\alpha f(x) = \partial^kf/\partial x_{1}^{\alpha_1} \cdots \partial x_{d}^{\alpha_d}$.
+This problem is in general well-posed whenever $2k > d$.
+In this regime, assuming that the $k$th order partial derivatives $D^\alpha f_0$ are all $\Leb^2(\Xset)$ bounded, the degree $k$ thin-plate spline has error on the order of \smash{$n^{-2k/(2k + d)}$} \citep{vandergeer2000}, which is minimax rate-optimal for such functions.
+Of course, assuming $f_0$ has $k$ bounded derivatives for some $2k > d$ is a very strong condition, but at present we do not know if (adaptations of) Laplacian smoothing on neighborhood graphs achieve these rates.
+
+\paragraph{Notation.}
+
+For an integer $p \geq 1$, we use $\Leb^p(\Xset)$ for the set of functions $f$ such that 
+\begin{equation*}
+\norm{f}_{\Leb^p(\Xset)}^p := \int_{\Xset} |f(x)|^p \,dx < \infty,
+\end{equation*}
+and $C^p(\Xset)$ for the set of functions that are $p$ times continuously differentiable.
+For sequences $a_n,b_n$, we write $a_n \lesssim b_n$ to mean $a_n \leq Cb_n$ for a constant $C>0$ and large enough $n$, and $a_n \asymp b_n$ to mean $a_n \lesssim b_n$ and $b_n \lesssim a_n$.
+Lastly, we use $a \wedge b = \max\{a,b\}$.
+
+\section{BACKGROUND}
+\label{sec:problem_setup_and_background}
+
+Before we present our main results in Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we define neighborhood graph Laplacians, and review known minimax rates over first-order Sobolev spaces.
+
+\paragraph{Neighborhood Graph Laplacians.}
+
+In the graph-based approach to nonparametric regression, we first build a neighborhood graph \smash{$G_{n,r} = (V,W)$}, for $V=\{1,\ldots,n\}$, to capture the geometry of $P$ (the design distribution) and $\Xset$ (the domain) in a suitable sense.
+The $n \times n$ weight matrix $W = (W_{ij})$ encodes proximity between pairs of design points; for a kernel function $K: [0,\infty) \to \Reals$ and radius $r > 0$, we have
 \begin{equation*}
 \label{eqn:neighborhood_graph}
-	{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr).
+W_{ij} = K\Biggl(\frac{\|X_i - X_j\|_2}{r}\Biggr),
 \end{equation*}
-	Then the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$, and the graph Laplacian can be written as
+with $\|\cdot\|_2$ denoting the $\ell_2$ norm on $\Rd$.
+Defining $D$ as the $n \times n$ diagonal matrix with entries \smash{$D_{ii} = \sum_{j = 1}^{n} W_{ij}$}, the graph Laplacian can then be written as
 \begin{equation}
 \label{eqn:graph_Laplacian}
-	\Lap_{n,r} = \bf{D} - \bf{W}
+\Lap_{n,r} = D - W.
 \end{equation}
-	\item The Laplacian smoothing estimator $\wh{f}_{\LS}$ \citep{smola2003} is a penalized least squares estimator, given by
+We add subscripts $n,r$ to $\Lap$ and $G$ to emphasize that they depend on the design points $X_1,\ldots,X_n$ and the radius $r$.
+To avoid any confusion: henceforth when we speak of \smash{$\wh{f}$} and \smash{$\wh{T}$}, we will always assume that they are fit using the neighborhood graph Laplacian \smash{$L = L_{n,r}$}.
+Writing its eigendecomposition as \smash{$L = \sum_{k = 1}^{n} \lambda_k v_k v_k^\top$}, we always assume, by convention, ordered eigenvalues $0 = \lambda_1 \leq \cdots \leq \lambda_n$, and unit-norm eigenvectors.
+%we will make reference to these objects later on.
+
+\paragraph{Sobolev Spaces.}
+
+We step away from graph-based methods for a moment, to briefly recall some classical results regarding minimax rates over Sobolev classes.
+We say that a function $f \in \Leb^2(\Xset)$ belongs to the \emph{first-order Sobolev space} $H^1(\Xset)$ if, for each $j = 1,\ldots,d$, the weak partial derivative $D^j f$ exists and belongs to $\Leb^2(\Xset)$.
+For such functions $f \in H^1(\Xset)$, the Sobolev seminorm \smash{$\seminorm{f}_{H^{1}(\Xset)}$} is the average size of the gradient $\nabla f = (D^1 f, \ldots, D^d f)$,
+\begin{equation*}
+\seminorm{f}_{H^1(\Xset)}^2 := \int_{\Xset} \bigl\|\nabla f(x)\bigr\|_2^2 \,dx,
+\end{equation*}
+with corresponding Sobolev norm 
+\begin{equation*}
+\norm{f}_{H^1(\Xset)} := \norm{f}_{\Leb^2(\Xset)} + |f|_{H^1(\Xset)}.
+\end{equation*}
+The Sobolev ball $H^1(\Xset,M)$ for $M > 0$ is
+\begin{equation*}
+H^1(\Xset,M) := \Bigl\{f \in H^1(\Xset): \norm{f}_{H^1(\Xset)}^2 \leq M^2\Bigr\}.
+\end{equation*}
+For further details regarding Sobolev spaces see, e.g., \citet{evans10,leoni2017}.
+
+\paragraph{Minimax Rates.}
+
+To carry out a minimax analysis of regression in Sobolev spaces, one must impose regularity conditions on the design distribution $P$. We shall assume the following.
+\begin{enumerate}[label=(P\arabic*)]
+	\item
+	\label{asmp:domain}
+	$P$ is supported on a domain $\Xset \subseteq \Rd$, which is an open, connected set with Lipschitz boundary.
+	\item
+	\label{asmp:density} 
+	$P$ admits a density $p$ such that
+	\begin{equation*}
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty, ~~\textrm{for all $x \in \Xset$}.
+	\end{equation*}
+	Additionally, $p$ is Lipschitz on $\Xset$, with Lipschitz constant $L_p$.
+\end{enumerate}
+
+Under conditions~\ref{asmp:domain}, \ref{asmp:density}, the minimax estimation rate over a Sobolev ball of radius $M \geq n^{-1/2}$ is (e.g., \citet{tsybakov2008_book}):
 \begin{equation}
-	\label{eqn:laplacian_smoothing}
-	\wh{f}_{\LS} := \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_{n,r} f \biggr\}
+\label{eqn:sobolev_space_estimation_minimax_rate}
+\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{2d/(2 + d)} n^{-2/(2 + d)}.
 \end{equation}
-	where $\rho > 0$ acts a tuning parameter on the penalty $f^T \Lap_{{n,r}} f$.
-	Assuming~\eqref{eqn:laplacian_smoothing} is a reasonable estimator of $f_0$, the squared empirical norm
+% RJT: This is a bit of a weird reference, because (as per my understanding) Tsybakov only proves the rate is tight for Holder classes, which are contained in the Sobolev class.
+So technically from Tsybakov alone we only know \gtrsim, not \asymp for this one.
+It's fine, I wouldn't change anything because it would involve explaining a lot more, I'm just pointing this out.
+(Throughout we assume $M \geq n^{-1/2}$, as otherwise the minimax rate is just the parametric rate $n^{-1}$.)
+
+As nonparametric hypothesis testing is (comparatively) less familiar than nonparametric estimation, we briefly summarize the main idea before stating the optimal error rate.
+In the goodness-of-fit testing problem, we ask for a test function---formally, a Borel measurable function $\phi$ taking values in $\{0,1\}$---which can distinguish between the hypotheses
 \begin{equation}
-	T_{\LS} = \frac{1}{n}\Bigl\|\wt{f}_{\LS}\Bigr\|_2^2 \label{eqn:laplacian_smoothing_test}
+\mathbf{H}_0: f_0 = f_0^{\star}, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{f_0^{\star}\}.
 \end{equation} 
-	is in turn a reasonable statistic to assess whether or not $f_0 = 0$. 
+Typically, the null hypothesis $f_0 = f_0^{\star} \in \mc{F}$ reflects the absence of interesting structure, and $\mc{F} \setminus  \{f_0^{\star}\}$ is a set of smooth departures from this null.
+In this paper, as in \citet{ingster2009}, we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\star} = 0$ and $\mc{F} = H^1(\Xset,M)$ is a first-order Sobolev ball.
+
+The minimax critical radius is the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{F}_{\epsilon} := \mc{F} \cap \{f: \|f\|_{\Leb^2(\Xset)} \geq \epsilon\}$:
+\begin{equation*}
+\epsilon(\mc{F}) := \inf\biggl\{\epsilon > 0: \inf_{\phi} \sup_{f_0 \in \mc{F}_{\epsilon}} \Ebb_{f_0}[1 - \phi] \leq \alpha \biggr\},
+\end{equation*} 
+where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Of course, the minimax critical radius $\epsilon$ depends on $\alpha$.
+However, we adopt the typical convention of treating $\alpha$ as a small but fixed positive constant; hence it will not affect the testing error rates, and we suppress it notationally.} See \citet{ingster82,ingster87,ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
+
+Testing $f_0=0$ is an easier problem than estimating $f_0$, and hence the minimax testing critical radius over $H^1(\Xset,M)$ is smaller than the minimax estimation rate, for $1 \leq d < 4$ (see \citet{ingster2009}):
+\begin{equation}
+\label{eqn:sobolev_space_testing_critical_radius}
+\epsilon^2\bigl(H^1(\Xset,M)\bigr) \asymp M^{2d/(4 + d)}n^{-4/(4 + d)}.
+\end{equation}
+When $d \geq 4$ the functions in $H^1(\Xset)$ are very irregular; formally speaking $H^1(\Xset)$ does not continuously embed into $\Leb^4(\Xset)$ when $d \geq 4$, and the minimax testing rates in this regime are unknown.
+
+\section{MINIMAX OPTIMALITY OF LAPLACIAN SMOOTHING}
+\label{sec:minimax_optimal_laplacian_smoothing}
+
+We now formalize the main conclusions of this paper: that Laplacian smoothing methods on neighborhood graphs are minimax rate-optimal over first-order continuum Sobolev classes.
+We will assume \ref{asmp:domain}, \ref{asmp:density} on $P$, and the following condition on the kernel $K$.
+\begin{enumerate}[label=(K\arabic*)]
+	\item
+	\label{asmp:kernel}
+	$K:[0,\infty) \to [0,\infty)$ is a nonincreasing function supported on $[0,1]$, its restriction to $[0,1]$ is Lipschitz, and $K(1) > 0$. Additionally, it is normalized so that
+	\begin{equation*}
+	\int_{\Rd} K(\|z\|_2) \,dz = 1.
+	\end{equation*}
+	We assume \smash{$\sigma_K = \frac{1}{d} \int_{\Rd} \|x\|_2^2 K(\|x\|_2) \,dx < \infty$}.
+\end{enumerate}
+This is a mild condition: recall the choice of kernel is under the control of the user, and moreover \ref{asmp:kernel} covers many common kernel choices.
+
+\paragraph{Estimation Error of Laplacian Smoothing.}
+
+Under these conditions, the Laplacian smoothing estimator \smash{$\wh{f}$} achieves an error rate that matches the minimax lower bound over $H^1(\Xset,M)$.
+This statement will hold whenever the graph $G_{n,r}$ is computed with radius $r$ in the following range.
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{0}
+	\item 
+	\label{asmp:ls_kernel_radius_estimation}
+	For constants $C_0,c_0>0$, the neighborhood graph radius $r$ satisfies
+	\begin{equation*}
+    C_0 \biggl(\frac{\log n}{n}\biggr)^{\frac{1}{d}} \leq r \leq c_0 \wedge M^{\frac{d - 4}{4 + 2d}} n^{-\frac{3}{4 + 2d}}.  
+	\end{equation*}
+\end{enumerate}
+
+Next we state Theorem~\ref{thm:laplacian_smoothing_estimation1}, our main estimation result. Its proof, as with all proofs of results in this paper, can be found in the supplementary document. 
+\begin{theorem}
+	\label{thm:laplacian_smoothing_estimation1}
+  Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ has dimension $d < 4$.
+Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume the neighborhood graph \smash{$G_{n,r}$} is computed with a kernel $K$ satifsying \ref{asmp:kernel}.
+There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_estimation}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$}, satisfies
+	\begin{equation*}
+	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(2 + d)} n^{-2/(2 + d)},
+	\end{equation*}
+  with probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c \min\{M^{d/(2d + 4)} n^{d/(2+d)},n\})$.
+% RJT: Alden, please verify that what I wrote is correct (I reorganized the way we explained this, added for $n \geq N$ ...).
+\end{theorem}
+
+To summarize: for $d = 1,2,$ or $3$, with high probability, the Laplacian smoothing estimator \smash{$\wh{f}$} has in-sample mean squared error that is within a constant factor of the minimax error.
+Some remarks:
+\begin{itemize}
+  \item The first-order Sobolev space $H^1(\Xset)$ does not continuously embed into $C^0(\Xset)$ when $d>1$ (in general, the $k$th order space $H^k(\Xset)$ does not continuously embed into $C^0(\Xset)$ except if $2k>d$).
+For this reason, one really cannot speak of pointwise evaluation of a Sobolev function $f_0 \in H^1(\Xset)$ when $d>1$ (as we do in Theorem~\ref{thm:laplacian_smoothing_estimation1} by defining our target of estimation to be $f_0(X_i)$, $i=1,\ldots,n$).
+We can resolve this by appealing to what are known as \emph{Lebesgue points}, as explained in the supplement.
+% RJT: Alden, please add a section to the supplement to explain this part.
+	\item The lower bound on $r$ imposed in condition \ref{asmp:ls_kernel_radius_estimation} is compatible with practice, where by far the most common choice of radius is the connectivity threshold \smash{$r \asymp (\log(n)/n)^{1/d}$}, which makes \smash{$G_{n,r}$} as sparse as possible while still being connected, for maximum computational efficiency.
+The upper bound may seem a bit more mysterious---we need it for technical reasons to ensure that \smash{$\wh{f}$} does not overfit, but note that as a practical matter one rarely chooses $r$ to be so large anyway.
+	\item It is possible to extend \smash{$\wh{f}$} to be defined on all of $\Xset$ and then evaluate the error of such an extension (as measured against $f_0$) in $\Leb^2(\Xset)$ norm.
+When \smash{$\wh{f}$} and $f_0$ are suitably smooth, tools from empirical process theory (see e.g., Chapter 14 of \citet{wainwright2019}) or approximation theory (e.g., Section 15.5 of \citet{johnstone2011}) guarantee that the $\Leb^2(\Xset)$ error is not too much greater than its in-sample counterpart.
+In fact, as shown in the supplement, if $f_0$ is Lipschitz smooth and we extend \smash{$\wh{f}$} to be piecewise constant over the Voronoi tesselation induced by $X_1,\ldots,X_n$, then the out-of-sample error \smash{$\|\wh{f}-f_0\|_{\Leb^2(\Xset)}$} is within a negligible factor of the in-sample error \smash{$\|\wh{f}-f_0\|_n$}.
+We leave analysis of the Sobolev case to future work.
+%	\item Under the stronger assumption that $f_0$ is Lipschitz smooth, we can replace the factor of $\delta$ in the high probability bound by the sharper $\delta^2/n$.
 \end{itemize}
 
-\textbf{(2.a)} Laplacian smoothing has many advantages compared to other nonparametric regression methods. For instance:
+When $d = 4$, our analysis results in an upper bound for the error of Laplacian smoothing that is within a \smash{$(\log n)^{1/3}$} factor of the minimax error rate.
+But when $d \geq 5$, our upper bounds are much worse.
+\begin{theorem}
+	\label{thm:laplacian_smoothing_estimation2}
+  Under the assumptions of Theorem~\ref{thm:laplacian_smoothing_estimation1}, if $\Xset$ has dimension $d = 4$ and \smash{$r \asymp (\log n/n)^{1/4}$}, then we obtain 
+	\begin{equation*}
+	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{4/3} \biggl(\frac{\log n}{n}\biggr)^{1/3},
+	\end{equation*}
+with the same probability guarantee as in Theorem~\ref{thm:laplacian_smoothing_estimation1}. If the dimension $\Xset$ is $d \geq 5$, TODO
+% RJT: Alden, please verify what I wrote is correct, and fill in the result for d \geq 5.
+\end{theorem}
 
+ This mirrors the conclusions of \citet{sadhanala16} who investigate estimation rates of Laplacian smoothing over the $d$-dimensional grid graph.
+These authors argue that their analysis is tight, and that it is likely the estimator, not the analysis, that is deficient when $d \geq 5$.
+Formalizing such a claim turns out to be harder in the random design setting than in the fixed design setting, and we leave it for future work.
+
+However, we do investigate the matter empirically.
+In Figure~\ref{fig:fig1}, we study the (in-sample) mean squared error of the Laplacian smoothing estimator as the dimension $d$ grows.
+Here $X_1,\ldots,X_n$ are sampled uniformly over $\Xset = [-1,1]^d$, and the regression function is taken as \smash{$f_0(x) \propto \Pi_{i = 1}^{d} \cos(a \pi x_i)$}, where $a = 2$ for $d = 2$, and $a = 1$ for $d \geq 3$.
+This regression function $f_0$ is quite smooth, and for $d = 2$ and $d = 3$ Laplacian smoothing appears to achieve or exceed the minimax rate.
+When $d = 4$, Laplacian smoothing appears modestly suboptimal; this fits with our theoretical upper bound, which includes a \smash{$(\log n)^{1/3}$} factor that plays a non-negligible role for these problem sizes ($n=1000$ to $n=10000$).
+On the other hand, when $d = 5$, Laplacian smoothing seems to be decidedly suboptimal.
+
+\begin{figure*}[tb]
+	\includegraphics[width=.245\textwidth]{figures/cosine/mse_by_sample_size_2d.pdf}
+	\includegraphics[width=.245\textwidth]{figures/cosine/mse_by_sample_size_3d.pdf} 
+	\includegraphics[width=.245\textwidth]{figures/cosine/mse_by_sample_size_4d.pdf}
+	\includegraphics[width=.245\textwidth]{figures/cosine/mse_by_sample_size_5d.pdf}
+	\caption{Mean squared error of Laplacian smoothing (\texttt{LS}) as a function of sample size $n$.
+Each plot is on the log-log scale, and the results are averaged over 5 repetitions, with Laplacian smoothing tuned for optimal average mean squared error.
+The black line shows the minimax rate (in slope only; the intercept is chosen to match the observed error).}
+	\label{fig:fig1}
+% RJT: Alden, please verify that what I wrote about the intercept above is right.
+And if so, can you redo the black line for d = 2 so that the intercept matches better?
+Thanks.
+Oh and if you're already going to redo one of these plots, you can satisfy my unreasonably nitpicky brain and change the x- and y-axes on all of them to use words rather than symbols/acronyms: "Sample size" and "Mean squared error".
+\end{figure*}
+
+\paragraph{Testing Error of Laplacian Smoothing.}
+
+For $b \geq 1$, define a threshold \smash{$\wh{t}_b$} as
+\begin{equation*}
+\wh{t}_{b} = \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}},
+\end{equation*}
+where we recall $\lambda_k$ is the $k$th smallest eigenvalue of \smash{$\Lap_{n,r}$}. The Laplacian smoothing test is then simply
+\begin{equation*}
+\wh{\varphi} = \1\bigl\{\wh{T} \leq \wh{t}_b\bigr\}.
+\end{equation*} 
+%(Clearly \smash{$\wh{\varphi}$} depends on $b$, but we suppress this notationally.) 
+Here $b \geq 1$ is a tuning parameter selected by the user, with the choice driven by the tolerated Type I versus Type II error.
+In fact, the threshold \smash{$\wh{t}_b$} is precisely the right choice to control the Type I error of \smash{$\wh{\varphi}$}: we show in the supplement that
+\begin{equation}
+\label{eqn:type_I_error}
+\Ebb_0\bigl[\wh{\varphi}\bigr] \leq \frac{1}{b^2}.
+\end{equation}
+The next theorem shows that this test also has small Type II error, uniformly over all $f_0$ that are separated from $0$ by at least the critical radius given in \eqref{eqn:sobolev_space_testing_critical_radius}.
+For this to hold, we will require a tighter range of scalings for the radius $r$.
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{1}
+	\item 
+	\label{asmp:ls_kernel_radius_testing}
+	For constants $C_0,c_0>0$, the neighborhood graph radius $r$ satisfies
+	\begin{equation*}
+	C_0\biggl(\frac{\log n}{n}\biggr)^{\frac{1}{d}} \leq r \leq c_0 \wedge M^{\frac{(d - 8)}{8 + 2d}} n^{\frac{d - 20}{32 + 8d}}.
+	\end{equation*}
+\end{enumerate}
+
+Now define
+\begin{equation*}
+M_{\max}(d) =
+\begin{cases*}
+n^{1/8} & \textrm{$d = 1$} \\
+n^{(4 - d)/(4d)} & \textrm{$d \geq 2$}.
+\end{cases*}
+\end{equation*}
+
+Note that when $d < 4$ and $M \leq M_{\max}(d)$, there always exists a choice of $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}. 
+We now state Theorem~\ref{thm:laplacian_smoothing_testing}, our main testing result.
+\begin{theorem}
+	\label{thm:laplacian_smoothing_testing}
+  Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ with $d < 4$, and where $M \leq M_{\max}(d)$.
+Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume \smash{$G_{n,r}$} is computed with a kernel $K$ satifsying \ref{asmp:kernel}.
+There are constants $N,C,c>0$ such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_testing}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$}, satisfies the following: for all $b \geq 1$, and all $f_0$ such that
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_testing}
+	\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2d/(4 + d)} n^{-4/(4 + d)},
+	\end{equation} 
+	the Type II error is upper bounded by
+	\begin{multline*}
+	\Ebb_{f_0}\bigl[1 - \wh{\varphi} \bigr] \leq \\
+	C\biggl(\frac{1}{b}\Bigl[1 + n^{-d/(4 + d)}M^{-2d/(4 + d)}\Bigr] + n\exp\bigl(-cnr^d\bigr)\biggr).
+	\end{multline*}
+\end{theorem}
+
+Some remarks:
 \begin{itemize}
-	\item It is fast, easy, and stable to compute.
-Letting ${\bf Y} = (Y_1,\ldots,Y_n) \in \Reals^n$, it is not hard to see that the minimizer $\wt{f}_{\LS} = (\rho \Lap_{n,r} + I)^{-1}{\bf Y}$ can be computed by solving a linear equation; in practice, one typically chooses $r$ to be small, with the result being that the matrix $\Lap_{n,r}$ is sparse.
-There exist known fast solvers of exactly this system.
-	\item Is is generalizable to non-standard data---for instance, text or images, or really any data modality on which it is possible to define a kernel.
-	\item It is easily adaptable to the semi-supervised learning setting. 
+	\item As mentioned earlier, Sobolev balls $H^1(\Xset,M)$ for $d \geq 4$ include quite irregular functions $f \not\in \Leb^4(\Xset)$.
+Proving tight lower bounds in this case is nontrivial, and as far as we understand such an analysis remains outstanding.
+On the other hand, if we explicitly assume that $f_0 \in \Leb^4(\Xset,M)$, then \citet{guerre02} show that the testing problem is characterized by a dimension-free lower bound $\epsilon^{2}(\Leb^4(\Xset,M)) \gtrsim n^{-1/2}$.
+Moreover, by training \smash{$\wh{f}$} to the interpolation limit, setting $\rho = 0$, the subsequent test \smash{$\wh{\varphi}$} will achieve (up to constants) this lower bound.
+That is, for any $f_0 \in \Leb^4(\Xset,M)$ such that \smash{$\|f_0\|_{\Leb^2(\Xset)}^2 \geq C b^2n^{-1/2}$}, we have both
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_testing_low_smoothness}
+	\Ebb_{f_0}\bigl[1 - \wh{\varphi}\bigr] \leq \frac{C(1 + M^4)}{b^2},
+	\end{equation} 
+	and \smash{$\Ebb_0[\wh{\varphi}] \leq 1/b^2$}.
+Note that the same results apply to \smash{$\wt{T}$}, since the thin-plate spline estimator \smash{$\wt{f}$} interpolates the responses $Y_1,\ldots,Y_n$ for $d>1$.
+	\item To compute the data-dependent threshold \smash{$\wh{t}_b$}, one must know all of the eigenvalues $\lambda_1,\ldots,\lambda_n$.
+Computing all these eigenvalues is far more expensive (cubic-time) than computing \smash{$\wh{T}$} in the first place (nearly-linear-time).
+But in practice we would not recommend using \smash{$\wh{t}_b$} anyway, and would instead we make the standard recommendation to calibrate via a permutation test \citep{hoeffding1952}.
 \end{itemize}
 
-\textbf{(2.b)} For this reason a considerable body of work has emerged analyzing the consistency of the graph Laplacian $\Lap_{{n,r}}$ as $n \to \infty$ and $r(n) \to 0$.
-This is meant in various senses:
-\begin{itemize}
-	\item Pointwise consistency, meaning $\Lap_{n,r}f \to \Delta_Pf$ for an appropriate limiting operator $\Delta_P$.
-	\item Spectral consistency, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$.
-	\item Consistency of norms, meaning $f^T \Lap_{n,f} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$. 
-\end{itemize}
-However, the statistical optimality of Laplacian smoothing is still not well understood.
+\paragraph{More Discussion of Variational Analog.}
 
-\textbf{(2.c)} Our main contributions lie in this latter direction, and they can all be summarized as follows: Laplacian smoothing methods are minimax optimal over Sobolev spaces.
-In more detail, we will show that when $f_0$ belongs to the Sobolev ball $H^1(\Xset;M)$:
-\begin{itemize}
-	\item \textbf{Minimax optimal estimation.} With high probability, $\frac{1}{n}\norm{\wt{f}_{\LS} - f_0}_{2}^2 \lesssim n^{-2/(2 + d)}$.
-	\item \textbf{Minimax optimal testing.}
-	A level-$\alpha$ test constructed using $T_{\LS}$ has non-trivial power whenever $\norm{f_0}_{\Leb^2(\Xset)}^2 \gtrsim n^{-4/(4 + d)}$. 
-	\item \textbf{Manifold adaptivity.}
-	If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$, both of the aforementioned rates hold with $d$ replaced by $m$.
-\end{itemize}
-Finally, we will also establish that the Laplacian smoothing estimator can be altered to take advantage of higher-order smoothness assumptions on $f_0$.
-This last conclusion will hold only under very limited circumstances.
+With some results in hand, let us pause to offer some explanation of why Laplacian smoothing can be optimal in settings where thin-plate splines are not even consistent.
+First, we elaborate on why this difference in performance is so surprising.
+As mentioned previously, the penalties in \eqref{eqn:laplacian_smoothing}, \eqref{eqn:thin_plate_spline} can be closely tied together: \citet{bousquet03} show that for $f \in C^2(\Xset)$,
+\begin{equation}
+\label{eqn:seminorm_consistency}
+\begin{aligned}
+\lim \frac{1}{n^2 r^{d + 2}} f^\top \Lap_{n,r} f & = \int_{\Xset} f(x) \cdot \Delta_Pf(x) p(x) \,dx \\
+& = \int_{\Xset} \|\nabla f(x)\|_2^2 p^2(x) \,dx.
+\end{aligned}
+\end{equation}
+In the above, the limit is as $n \to \infty$ and $r \to 0$, $\Delta_P$ is the (weighted) Laplace-Beltrami operator
+\begin{equation*}
+\Delta_Pf := -\frac{1}{p} \dive\bigl(p^2\nabla f),
+\end{equation*}
+and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies e.g., Dirichlet boundary conditions.} To be clear, this argument does not formally imply that $\wh{f}$ and $\wt{f}$ are close (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of continuous functions with weak derivatives in $\Leb^2(\Xset)$).
+But it does seem to suggest that the two estimators should behave somewhat similarly.
 
-\paragraph{Organization.}
+Of course, we know this is not the case: $\wh{f}$ and $\wt{f}$ look very differently when $d > 1$.
+What is driving this difference?
+The key point is that the discretization imposed by the graph $G_{n,r}$---which might seem problematic at first glance---turns out to be a blessing.
+The problem with~\eqref{eqn:thin_plate_spline} is that the class $H^1(\Xset)$, which fundamentally underlies the criterion, is far ``too big'' for $d > 1$.
+This is meant in various related senses.
+By the Sobolev embedding theorem, for $d>1$, the class $H^1(\Xset)$ does not continuously embed into any H\"{o}lder space; and in fact it does not even continuously embed into $C^0(\Xset)$.
+Thus we cannot really restrict the optimization to \emph{continuous} and weakly differentiable functions, as we could when $d=1$ (the smoothing spline case), without throwing out a substantial subset of functions in $H^1(\Xset)$.
+Even among continuous and differentiable functions $f$, as we explained previously, we can use ``bump'' functions (as in \citet{green93}) to construct $f$ that interpolates the pairs $(X_i,Y_i)$, $i=1,\ldots,n$ and achieves arbitrarily small penalty (and hence criterion) in \eqref{eqn:thin_plate_spline}.
+In this sense, any estimator resulting from solving \eqref{eqn:thin_plate_spline} estimator will clearly be inconsistent.
 
-List out what is to come.
+On the other hand, problem \eqref{eqn:laplacian_smoothing} is finite-dimensional.
+As a result \smash{$\wh{f}$} has far less capacity to overfit than does \smash{$\wt{f}$}, for any given sample size $n$.
+Discretization is not the only way to make the problem \eqref{eqn:thin_plate_spline} more tractable: for instance, one can replace the penalty \smash{$\int_{\Xset} \|\nabla f(x)\|_2^2 \,dx$} with a stricter choice like \smash{$\esssup_{x \in \Xset} \|\nabla f(x)\|_2$}, or conduct the optimization over some finite-dimensional linear subspace of $H^1(\Xset)$ (i.e., use a sieve).
+While these solutions do improve the statistical properties of \smash{$\wt{f}$} for $d > 1$ (see e.g., \citet{birge1993,birge1998,vandergeer2000}), Laplacian smoothing is generally speaking much simpler and more computationally friendly, and in addition, the other approaches are specifically tailored to the domain $\Xset$, in stark to \smash{$\wh{f}$}.
 
-\section{Minimax regression over Sobolev spaces}
+\paragraph{Overview of Analysis.}
 
-\textbf{(3)} Sobolev spaces are classical function spaces, over which linear nonparametric regression methods.
-One notable such method is smoothing splines, to which we give special attention because of their close connection to Laplacian smoothing.
+The comparison with thin-plate splines highlights some surprising differences between \smash{$\wh{f}$} and \smash{$\wt{f}$}.
+Such differences also preclude us from analyzing \smash{$\wh{f}$} by, say, using \eqref{eqn:seminorm_consistency} to establish a coupling between \smash{$\wh{f}$} and \smash{$\wt{f}$}---we know this cannot work, because we would like to prove meaningful error bounds on \smash{$\wh{f}$} in regimes where no such bounds exist for \smash{$\wt{f}$}.
 
-\textbf{(4)} In fact, an alternative way to motivate Laplacian smoothing nonparametric regression methods---as opposed to the advantages mentioned in \textbf{(2.a)}---comes from viewing them as discrete-time versions of smoothing splines.
-This is made more rigorous by the consistency results of \textbf{(2.b)}.
-Viewed in this light, an additional advantage of Laplacian smoothing emerges: it automatically adapts to the geometry of $\Xset$ and $p$.
+Instead we take a different approach, and directly analyze the error of \smash{$\wh{f}$} and \smash{$\wh{T}$} using a bias-variance decomposition (conditional on $X_1,\ldots,X_n$).
+A standard calculation shows that
+\begin{equation*}
+\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \underbrace{\vphantom{\sum_{k=1}^n}\frac{2\rho}{n} \bigl(f_0^\top \Lap_{n,r} f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{10}{n} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2}}_{\textrm{variance}},
+\end{equation*}
+and likewise that $\wh{\varphi}$ has high power whenever
+\begin{equation*}
+\bigl\|f_0\bigr\|_n^2 \geq \underbrace{\vphantom{\sqrt{\sum_{k=1}^n}}\frac{2\rho}{n} \bigl(f_0^\top \Lap_{n,r} f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{4b}{n} \sqrt{\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}}}_{\textrm{variance}}.
+\end{equation*}
+The bias and variance terms are each functions of the random graph \smash{$G_{n,r}$}, and hence are themselves random.
+To upper bound them, we build on some recent works \citep{burago2014,trillos2019,calder2019} regarding the consistency of neighborhood graphs to establish the following lemmas.
+These lemmas assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and \ref{asmp:kernel} on the kernel used to compute the neighborhood graph \smash{$G_{n,r}$}.
 
-\textbf{(5)} Smoothing splines are known to have strong minimax optimal properties. Reasoning by analogy, we might hope that Laplacian smoothing would share these properties.
+\begin{lemma}
+	\label{lem:graph_sobolev_seminorm}
+  There are constants $N,C_2 > 0$ such that for $n \geq N$, $r \leq c_0$, and $f \in H^1(\Xset)$, with probability at least $1 - \delta$, it holds that
+	\begin{equation}
+	\label{eqn:graph_sobolev_seminorm}
+	f^\top \Lap_{n,r} f \leq \frac{C_2}{\delta} n^2 r^{d + 2} |f|_{H^1(\Xset)}^2.
+	\end{equation}
+\end{lemma}
 
-\section{Graph Sobolev classes}
+\begin{lemma}
+	\label{lem:neighborhood_eigenvalue} 
+  There are constants $N,C_1,C_3,c_1,c_3 > 0$ such that for $n \geq N$, $C_0(\log n/n)^{1/d} \leq r \leq c_0$, and $f \in H^1(\Xset)$, with probability at least $1 - C_1n\exp(-c_1nr^d)$, it holds that
+	\begin{equation}
+	\label{eqn:neighborhood_eigenvalue}
+	c_3A_{n,r}(k) \leq \lambda_k \leq C_3A_{n,r}(k), ~~\textrm{for $2 \leq k \leq n$},
+	\end{equation}
+where \smash{$A_{n,r}(k) = \min\{nr^{d + 2}k^{2/d},nr^d\}$}.
+\end{lemma}
 
-\textbf{(6)} The minimax optimality of smoothing splines rests on two crucial facts: one, the a priori assumption that $|f|_{H^1(\Xset)}$ is bounded; two, an upper bound on the metric entropy of $H^1(\Xset)$, characterized by the decay of its eigenvalues.
-\textcolor{red}{(TODO)}:(I know this isn't well-stated, and I need to clean it up.)  We now show that qualitatively similar statements hold with respect to
+Lemma~\ref{lem:graph_sobolev_seminorm} gives a direct upper bound on the bias term.
+Lemma~\ref{lem:neighborhood_eigenvalue} leads to a sufficiently tight upper bound on the variance term whenever the radius $r$ is sufficiently small; precisely, when $r$ is upper bounded as in \ref{asmp:ls_kernel_radius_estimation} for estimation, or \ref{asmp:ls_kernel_radius_testing} for testing.
+The parameter $\rho$ is then tuned to minimize the sum of bias and variance, in the usual manner.
 
-\quad \textbf{(6.1)} the graph Sobolev semi-norm $f_0^T \Lap_{n,r} f_0$ and,
+It may be useful to give one more perspective on our approach.
+A common strategy in analyzing penalized least squares estimators is to assume two properties: first, that the regression function $f_0$ lies in (or near) a ball defined by the penalty operator; second, that this ball is reasonably small, e.g., as measured by metric entropy, or Rademacher complexity, etc.
+In contrast, in Laplacian smoothing, the penalty induces a ball
+\begin{equation*}
+H^1(G_{n,r},M) := \{f: f^\top \Lap_{n,r} f \leq M^2\}
+\end{equation*}
+that is data-dependent and random, and so we do not have access to either of the aforementioned properties a priori, and instead, must prove they hold with high probability.
+In this sense, our analysis is different than the typical one in nonparametric regression.
 
-\quad \textbf{(6.2)} the graph eigenvalues $\lambda_k(\Lap_{n,r})$.
+\section{MANIFOLD ADAPTIVITY}
+\label{sec:manifold_adaptivity}
 
-\section{Minimax optimality of Laplacian smoothing}
+The minimax rates $n^{-2/(2 + d)}$ and $n^{-4/(4 + d)}$, in estimation and testing, suffer from the curse of dimensionality.
+However, in practice it can be often reasonable to assume a \emph{manifold hypothesis}: that the data $X_1,\ldots,X_n$ lie on a manifold $\Xset$ of $\Rd$ that has intrinsic dimension $m < d$.
+Under such an assumption, it is known \citep{bickel2007,ariascastro2018} that the optimal rates over $H^1(\Xset)$ are now $n^{-2/(2 + m)}$ (for estimation) and $n^{-4/(4 + m)}$ (for testing), which are much faster than the full-dimensional error rates when $m \ll d$.
 
-\textbf{(7)} Using \textbf{(6.1)} and \textbf{(6.2)}, we establish upper bounds on the estimation and testing error of Laplacian smoothing, which confirm that Laplacian smoothing is asymptotically minimax rate-optimal over the Sobolev class $H^1(\Xset)$.
+% RJT: Again, would venture a guess that these are stated results for Holder classes, not Sobolev classes.
+So same comment as before, as when we referenced Tsybakov.
+Just checking my understanding ...
+not sure it's worth changing this at this point.
 
-\textbf{(8)} While we motivated graph Laplacians by suggesting they might replicate the statistical properties of smoothing splines, we have in fact managed to show something stronger. In particular:
+On the other hand, a theory has been developed \citep{belkin03,belkin05,niyogi2013} establishing that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+We contribute to this line of work by showing that under the manifold hypothesis, Laplacian smoothing achieves the sharper minimax rates over $H^1(\Xset)$.
 
-\quad \textbf{(8.1)} When $d \geq 2$, there is no Sobolev embedding of $H^1(\Xset) \subseteq C^s(\Xset)$ for any $s > 0$. Therefore the smoothing spline estimator is ill-posed.
+\paragraph{Error Rates Assuming the Manifold Hypothesis.}
 
-\quad \textbf{(8.2)} Additionally, the metric entropy of $C^1(\Xset)$ is too large to obtain optimal rates for \textcolor{red}{Holder-smoothing splines} when $d \geq 2$.
-Thus, we see a gap emerge between the optimality properties of the discrete-time (Laplacian smoothing) and continuous-time (smoothing splines) estimators, in favor of the latter.
+The conditions and results presented here will be largely similar to the previous ones, except with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
+For the remainder, we assume the following.
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{2}
+	\item 
+	\label{asmp:domain_manifold}
+	$P$ is supported on a compact, connected, smooth manifold $\Xset$ embedded in $\Rd$, of dimension $m \leq d$. The manifold is without boundary and has positive reach \citep{federer1959}.
+	\item 
+	\label{asmp:density_manifold} 
+  $P$ admits a density $p$ with respect to the volume form of $\Xset$ such that 
+	\begin{equation*}
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty, ~~\textrm{for all $x \in \Xset$}. 
+	\end{equation*}
+	Additionally, $p$ is Lipschitz on $\Xset$, with Lipschitz constant $L_p$.
+\end{enumerate}
 
-\quad \textbf{(8.3)} On the other hand, when $d \geq 4$, we no longer get the known optimal rates in the estimation problem.
-When $d = 4$, our rates are suboptimal by a factor of $\log n$.
-When $d > 5$ they are very suboptimal.
-We believe this is related to classical results regarding the entropy of low-smoothness Holder classes, i.e.
-the results used to justify \textbf{(8.2)}.
-The interesting point is that the parameter spaces $C^1(\Xset)$ and $H^1(G_{n,r})$ become ``too large'' for different values of $d$.
+Under the assumptions \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}, and \ref{asmp:kernel}, and for a suitable range of $r$, the error bounds on the estimator \smash{$\wh{f}$} and test \smash{$\wh{\varphi}$} will depend on $m$ instead of $d$.
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{3}
+	\item 
+	\label{asmp:ls_kernel_radius_estimation_manifold}
+	For constants $C_0,c_0>0$, the neighborhood graph radius $r$ satisfies 
+	\begin{equation*}
+	C_0\biggl(\frac{\log n}{n}\biggr)^{\frac{1}{m}} \leq r \leq c_0 \wedge  M^{\frac{(m - 4)}{(4 + 2m)}} n^{\frac{-3}{(4 + 2m)}}.
+	\end{equation*}
+\end{enumerate}
 
-\quad \textbf{(8.4)} \textcolor{red}{(TODO)}: Translate these results into error in $\Leb^2(P)$ error, if possible.
+\begin{theorem}
+	\label{thm:laplacian_smoothing_estimation_manifold}
+  As in Theorem \ref{thm:laplacian_smoothing_estimation1}, but where $\Xset \subseteq \Rd$ is now a manifold with intrinsic dimension $m < 4$, and the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}.
+There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_estimation_manifold}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + m)} (nr^{m + 2})^{-1} n^{-2/(2 + m)}$}, satisfies
+	\begin{equation*}
+	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{2m/(2 + m)} n^{-2/(2 + m)},
+	\end{equation*}
+  with probability at least $1 - \delta -  C_1 n\exp(-cnr^m) - \exp(-c \min\{M^{m/(2m + 4)} n^{m/(2+m)},n\})$.
+% RJT: Alden, please verify that what I wrote is correct.
+\end{theorem}
 
-\subsection{Manifold adaptivity}
+\begin{enumerate}[label=(R\arabic*)]
+	\setcounter{enumi}{4}
+	\item 
+	\label{asmp:ls_kernel_radius_testing_manifold}
+	For constants $C_0,c_0>0$, the neighborhood graph radius $r$ satisfies 
+	\begin{equation*}
+	C_0\biggl(\frac{\log n}{n}\biggr)^{\frac{1}{m}} \leq r \leq c_0 \wedge M^{\frac{(m - 8)}{8 + 2m}} n^{\frac{m - 20}{32 + 8m}}.
+	\end{equation*}
+\end{enumerate}
 
-\quad \textbf{(9)} We now show that when $\Xset$ is an $m$-dimensional submanifold of $\Reals^d$, Laplacian smoothing approaches automatically ``feel'' the intrinsic dimension $m$ of $\Xset$, which translate into improved upper bounds on estimating and testing error.
-This confirms the statement in \textbf{(4)} that Laplacian smoothing automatically adapts to the geometry of $\Xset$ and $p$.
+\begin{theorem}
+	\label{thm:laplacian_smoothing_testing_manifold}
+  As in Theorem \ref{thm:laplacian_smoothing_testing}, but where $\Xset \subseteq \Rd$ is a manifold with intrinsic dimension $m < 4$, $M \leq M_{\max}(m)$, and the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}.
+There are constants $N,C,c>0$ such that for any $n \geq N$, and any $r$ as in \ref{asmp:ls_kernel_radius_testing_manifold}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$}, satisfies the following: for all $b \geq 1$, and $f_0$ such that
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_testing_manifold}
+	\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2m/(4 + m)} n^{-4/(4 + m)},
+	\end{equation} 
+	the Type II error is upper bounded by
+	\begin{multline*}
+	\Ebb_{f_0}\bigl[1 - \wh{\varphi} \bigr] \leq \\
+	C\biggl(\frac{1}{b}\Bigl[1 + n^{-m/(4 + m)}M^{-2m/(4 + m)}\Bigr] + n\exp\bigl(-cnr^m\bigr)\biggr).
+	\end{multline*}
+% RJT: Alden, please verify that what I wrote is correct.
+\end{theorem}
 
-\subsection{Higher-order smoothness classes}
+The proof of Theorems \ref{thm:laplacian_smoothing_estimation_manifold} and \ref{thm:laplacian_smoothing_testing_manifold} proceeds in a similar manner to that of Theorems \ref{thm:laplacian_smoothing_estimation1} and \ref{thm:laplacian_smoothing_testing}.
+The key difference is that in the manifold setting, the equations \eqref{eqn:graph_sobolev_seminorm} and \eqref{eqn:neighborhood_eigenvalue} used to upper bound bias and variance will hold with $d$ replaced by $m$.
 
-\quad \textbf{(10)} Analogously to smoothing splines, Laplacian smoothing can also be adapted to take advantage of additional assumed regularity on $f_0$, i.e.
-$f_0 \in H^s(\mc{\Xset})$ for $s > 1$.
-In very limited circumstances, we can show that this adapted estimator achieves the sharper minimax rates over these classes.
-However, the general story--for all combinations of $s$ and $d$---remains beyond our reach.
+We emphasize that little about $\Xset$ needs be known for Theorems~\ref{thm:laplacian_smoothing_estimation_manifold} and \ref{thm:laplacian_smoothing_testing_manifold} to hold.
+Indeed, all that is needed is the intrinsic dimension $m$, to properly tune $r$ and $\rho$ (from a theoretical point of view), and otherwise \smash{$\wh{f}$} and \smash{$\wh{\varphi}$} are computed without regard to $\Xset$.
+In contrast, the penalty in \eqref{eqn:thin_plate_spline} would have to be specially tailored to work in this setting, revealing another advantage of the discrete approach to the variational one.
 
-\section{Simulations}
+\section{DISCUSSION}
+\label{sec:discussion}
 
-\quad \textbf{(11)} Empirically, we demonstrate that the risk for Laplacian smoothing is comparable to that of smoothing splines, in the full dimensional case.
-Although one cannot carry out a comparison in the case of an unknown manifold --- since the smoothing spline estimator must be changed in a non-obvious manner --- we do show that when the density $p$ is concentrated close to a manifold and $f_0$ is smooth where $p$ is peaked, we have improved performance for $\wh{f}_{\LS}$ compared to $\wh{f}_{\mathrm{SM}}$.
-
-\section{Discussion}
-
-\clearpage
+We have shown that Laplacian smoothing, computed over a neighborhood graph, can be optimal for both estimation and goodness-of-fit testing over Sobolev spaces.
+There are many extensions worth pursuing, and several have already been mentioned.
+We conclude by mentioning a couple more.
+In practice, it is more common to use $k$-nearest-neighbor (kNN) graphs than neighborhood graphs, due to the guaranteed connectivity and sparsity of the former; we suspect that by building on the work of \citet{calder2019}, one can show that our main results all hold under the kNN graph as well.
+In another direction, one can also generalize Laplacian smoothing by replacing the penalty \smash{$f^\top \Lap_{n,r} f$} with \smash{$f^\top \Lap_{n,r}^k f$} for $k > 1$.
+The hope is that this would then achieve minimax optimal rates over the higher-order Sobolev class $H^k(\Xset)$.
+% In the very special case of $k = 2$ and $d \leq 2$, we can show that this is indeed the case, but the general story--for all combinations of $k$ and $d$---remains beyond our reach.
 
 \bibliographystyle{plainnat}
-\bibliography{../../../graph_testing_bibliography}
+\bibliography{../../../graph_regression_bibliography}
 

### 2021-02-23 manuscripts/laplacian_smoothing/AISTATS/supplement_minimax_optimal_laplacian_smoothing.tex
Detailed pass over the main AISTATS paper

@@
 \usepackage{verbatim}
+\usepackage{microtype}
 
@@
 \end{equation*}
-To avoid confusion, in this section we write $\lambda_k(G_{n,r})$ for the $k$th smallest eigenvalue of the graph Laplacian matrix $\Lap_{n,r}$ and $\lambda_k(\Delta_P)$ for the $k$th smallest eigenvalue of $\Delta_P$ \footnote{Under the assumptions~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density}, the operator $\Delta_P$ has a discrete spectrum; see \citet{garciatrillos18} for more details.}.
+To avoid confusion, in this section we write $\lambda_k(G_{n,r})$ for the $k$th smallest eigenvalue of the graph Laplacian matrix $\Lap_{n,r}$ and $\lambda_k(\Delta_P)$ for the $k$th smallest eigenvalue of $\Delta_P$ \footnote{Under the assumptions~\ref{asmp:domain} and~\ref{asmp:density}, the operator $\Delta_P$ has a discrete spectrum; see \citet{garciatrillos18} for more details.}.
 Some other notation: throughout this section, we will write $A, A_0, A_1,\ldots$ and $a,a_0,a_1,\ldots$ for constants which may depend on $\Xset$, $d$, $K$, and $p$, but do not depend on $n$; we keep track of all such constants explicitly in our proofs.
@@
 	\label{prop:weyl}
-	Suppose the density $p$ and the domain $\Xset$ satisfy~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density}. Then there exist constants $a_2$ and $A_2$ such that
+	Suppose the density $p$ and the domain $\Xset$ satisfy~\ref{asmp:domain} and~\ref{asmp:density}. Then there exist constants $a_2$ and $A_2$ such that
 	\begin{equation}
@@
 \end{proposition}
-See Lemma 28 of~\citet{dunlop2020} for a proof that~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density} imply Weyl's Law.
+See Lemma 28 of~\citet{dunlop2020} for a proof that~\ref{asmp:domain} and~\ref{asmp:density} imply Weyl's Law.
 \paragraph{Proof of Lemma~\ref{lem:neighborhood_eigenvalue}.}
@@
 Let us verify that $\lambda_{\ell_{\star}}(\Delta_P)$ satisfies the condition~\ref{eqn:neighborhood_eigenvalue_1} of Theorem~\ref{thm:neighborhood_eigenvalue}.
-Setting $c_0 := 1/(2^{1/d}4A_2^{1/2})$, the assumed upper bound on the radius $r < c_0$ in \ref{asmp:radius1} guarantees that $\ell_{\star} \geq 2$.
+Setting $c_0 := 1/(2^{1/d}4A_2^{1/2})$, the assumed upper bound on the radius $r \leq c_0$ guarantees that $\ell_{\star} \geq 2$.
 Therefore, by Proposition~\ref{prop:weyl} we have that
@@
 	\label{prop:optimal_transport}
-	Suppose $\Xset$ satisfies~\ref{asmp:domain} and $p$ satisfies~\ref{asmp:bounded_lipschitz_density}.
+	Suppose $\Xset$ satisfies~\ref{asmp:domain} and $p$ satisfies~\ref{asmp:density}.
 Then with probability at least $1 - A_0 n \exp\bigl\{-a_0 n\theta^2\wt{\delta}^d\bigr\}$, the following statement holds: there exists a probability measure $\wt{P}_n$ with density $\wt{p}_n$ such that:
@@
 	\label{lem:empirical_norm_sobolev}
-	Fix $\delta \in (0,1)$ Suppose $P$ satisfies~\ref{asmp:bounded_lipschitz_density}. If $f \in H^1(\Xset,M)$ is lower bounded in $\Leb^2(\Xset)$ norm,
+	Fix $\delta \in (0,1)$ Suppose $P$ satisfies~\ref{asmp:density}. If $f \in H^1(\Xset,M)$ is lower bounded in $\Leb^2(\Xset)$ norm,
 	\begin{equation}

### 2021-02-24 manuscripts/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
Style file and preamble

@@
 \usepackage[accepted]{aistats2021}
+
+\special{papersize = 8.5in, 11in}
+\setlength{\pdfpageheight}{11in}
+\setlength{\pdfpagewidth}{8.5in}
+
 \usepackage{amsmath, amsfonts, amsthm, amssymb}
@@
 	In this paper we study the statistical properties of Laplacian smoothing, a graph-based approach to nonparametric regression.
-Under standard regularity conditions, we establish upper bounds on the error of the Laplacian smoothing estimator \smash{$\wh{f}$}, and a goodness-of-fit test based on \smash{$\wh{f}$}.
+Under standard regularity conditions, we establish upper bounds on the error of the Laplacian smoothing estimator \smash{$\wh{f}$}, and a goodness-of-fit test also based on \smash{$\wh{f}$}.
 These upper bounds match the minimax optimal estimation and testing rates of convergence over the first-order Sobolev class $H^1(\Xset)$, for $\Xset \subseteq \Rd$ and $1 \leq d < 4$; in the estimation problem, for $d = 4$, they are optimal modulo a $\log n$ factor.

### 2021-02-24 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Added a note on some pitfalls to avoid when thinking about Sobolev spaces.

@@
 \textbf{(3) Why we study.} Laplacian eigenmaps is thus an example of a \emph{spectral series} estimator.
-Series estimators are one of the most classical ways to perform non-parametric regression.
-Traditionally, series estimators are defined with respect to a fixed set of basis functions, orthogonalized with respect to some pre-determined reference measure $Q$.
-In contrast, eigenvectors of the graph Laplacian are data-dependent objects, and adapt to the geometry of the distribution $P$ in a rich manner.
+A spectral series estimator is a special case of an orthogonal series estimator, one of the most classical methods of non-parametric regression.
+Traditionally, orthogonal series estimators are defined with respect to a fixed set of basis functions, orthogonalized with respect to some pre-determined reference measure $Q$.
+Spectral series estimators take a particular orthogonal basis, the eigenfunctions of $\Delta_Q$.
+In contrast, eigenvectors of the graph Laplacian are data-dependent objects, and adapt to the geometry of the design distribution $P$ in a rich manner.
 
-The study of (classical) spectral series estimators is rich, and their theoretical optimality is well-understood \textcolor{red}{(references)}.
+The theoretical properties of (classical) spectral series estimators are by this point well-understood \textcolor{red}{(references)}.
+In particular, they are minimax optimal for nonparametric regression over H\"{o}lder and Sobolev spaces.
 However, in practice these estimators suffer from some serious drawbacks.
+Finding the eigenfunctions of $\Delta_Q$ is in general non-trivial.
+The statistical optimality of such estimators also hinges on the reference measure $Q$ being equal to $P$, an unrealistic assumption when $P$ is unknown.
+If $Q \neq P$, the standard recommendation is to instead do least squares---this can be statistically rate-optimal, but is numerically unstable when dimension $d$ is even moderately large.
 
@@
 Broadly speaking, we show that when the regression function $f_0$ is smooth, in the sense of having bounded derivatives in the Sobolev sense, Laplacian eigenmaps is statistically minimax optimal.
-This statement holds for different relations between the dimension $d$ and number of derivatives $s$, depending on the problem (estimation) \textcolor{red}{(TODO): Add two tables summarizing your major results here.}
+This statement holds for different relations between the dimension $d$ and number of derivatives $s$, depending on the problem (estimation or testing).
+\textcolor{red}{(TODO): Add two tables summarizing your major results here.}
 
 \textbf{(5) Organization.}
+
+\textcolor{red}{(TODO)}
 
@@
 \begin{equation*}
-H^s(\mc{X};M) := \{f \in H^s(P): \|f\|_{H^s(P)} \leq M\}
+H^s(P;M) := \{f \in H^s(P): \|f\|_{H^s(P)} \leq M\}
 \end{equation*}
@@
 \label{sec:first_order_sobolev_classes}
-To begin, we handle the case where $f_0 \in H^1(\mc{X})$.
+To begin, we handle the case where $f_0 \in H^1(P)$.
+We show that $\wh{f}$ and a test based on $\wh{T}$ are minimax optimal, for all values of $d$, and under no additional assumptions on the data generating process, i.e.
+on $P$ or $f_0$.
 
-\textbf{(1) Laplacian eigenmaps is an optimal estimator.} When the graph radius $\varepsilon$ and number of eigenvectors $K$ are chosen appropriately, the estimator $\wh{f}$ has (in-sample) optimal risk (up to constant factors) $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2/(2 + d)}$.
+\paragraph{Estimation.} When the graph radius $\varepsilon$ and number of eigenvectors $K$ are chosen appropriately, the estimator $\wh{f}$ has (in-sample) optimal risk (up to constant factors) $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2/(2 + d)}$.
 \begin{enumerate}[label=(A\arabic*)]
@@
 	\item 
-	\label{asmp:estimation_fo} 
+	\label{asmp:parameters_estimation_fo} 
 	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy the following inequalities:
 	\begin{equation}\\
-	\label{eqn:radius_estimation_fo} 
-	C_1 \biggl(\frac{\log n}{n}\biggr)^{1/d} \leq \varepsilon \leq \min\{i_0,K^{-1/d}\},
+	\label{eqn:radius_fo} 
+	C_0 \biggl(\frac{\log n}{n}\biggr)^{1/d} \leq \varepsilon \leq \min\{i_0,K^{-1/d}\},
 	\end{equation}
@@
 	\label{eqn:eigenvector_estimation_fo} 
-	c_2 (M^2 n)^{2d/(2 + d)}\leq K \leq C_2 (M^2 n)^{2d/(2 + d)}.
+	c_1 (M^2 n)^{2d/(2 + d)}\leq K \leq C_1 (M^2 n)^{2d/(2 + d)}.
 	\end{equation}
@@
 	\label{thm:laplacian_eigenmaps_estimation_fo}
-	Suppose the general assumptions~\ref{asmp:domain}-\ref{asmp:sobolev_radius} are satisfied.
-Then for any $f_0 \in H^1(P,M)$, if Laplacian eigenmaps is run with parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:estimation_fo}, with probability at least $1 - \delta - \textcolor{red}{(???)}$ it follows that
+	Suppose $f_0 \in H^1(P,M)$.
+If the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_fo}, then the following statement holds for any $\delta \in (0,1)$: with probability at least $1 - \delta - \textcolor{red}{(???)}$,
 	\begin{equation}
@@
 \begin{itemize}
-	\item \textbf{(1.1) How to tune hyper-parameters.} We note that the ranges~\eqref{eqn:radius_estimation_fo} and~\eqref{eqn:eigenvector_estimation_fo} depend on unknown quantities such as $f_0$ and $P$.
+	\item \textbf{(1.1) How to tune hyper-parameters.} We note that the ranges~\eqref{eqn:radius_fo} and~\eqref{eqn:eigenvector_estimation_fo} depend on unknown quantities such as $f_0$ and $P$.
 In practice, one typically tunes hyper-parameters by sample-splitting or cross-validation.
@@
 The choice of $t_{\delta}$ guarantees that $\varphi$ is a level-$\delta$ test.
-Under appropriate choices of $\varepsilon$ and $K$, the test $\varphi$ has non-negligible power (up to constant factors) against alternatives separated the null by at least $\|f_0\|_{P}^2 \gtrsim M^{2d/(4 + d)}n^{-4/(4 + d)}$.
+Under appropriate choices of $\varepsilon$ and $K$, the test $\varphi$ has non-negligible power (up to constant factors) against alternatives separated the null by at least $\|f_0\|_{P}^2 \gtrsim M^{2d/(4 + d)}n^{-4/(4 + d)}$, whenever $d < 4$.
 
+\begin{enumerate}[label=(A\arabic*)]
+	\setcounter{enumi}{4}
+	\item 
+	\label{asmp:parameters_testing_fo}
+	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy~\eqref{eqn:radius_fo}. Additionally,
+	\begin{equation}
+	\label{eqn:eigenvector_testing_fo}
+	c_2 (M^2 n)^{2d/(4 + d)}\leq K \leq C_2 (M^2 n)^{2d/(4 + d)}.
+	\end{equation}
+\end{enumerate}
+Let $p(d) = \min\{1,2/d\}$.
 \begin{theorem}
 	\label{thm:laplacian_eigenmaps_testing_fo}
-	\textcolor{red}{(TODO)}
+	Suppose $f_0 \in H^1(\mc{X};M)$.
+If the Laplacian eigenmaps test $\varphi$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_testing_fo}, then the following statement holds for any $\delta \in (0,1)$: if $f_0$ satisfies
+	\begin{equation}
+	\label{eqn:laplacian_eigenmaps_testing_criticalradius_fo}
+	\|f_0\|_P^2 \geq \frac{C}{\delta} \max\biggl\{\frac{1}{\delta}M^{2d/(4 + d)}n^{-4/(4 + d)}, M n^{-p(d)}\biggr\},
+	\end{equation}
+	then $\varphi$ has Type II error upper bounded by 
+	\begin{equation}
+	\Pbb_{f_0}(\varphi = 0) \leq \delta + \textcolor{red}{(...)}
+	\end{equation}
 \end{theorem}

### 2021-02-24 notes/misc/avoiding_pitfalls_sobolev.tex
Added a note on some pitfalls to avoid when thinking about Sobolev spaces.

--- 2021-02-11 manuscripts/laplacian_eigenmaps/outline.tex
@@
 
+%%% Integrals
+\def\Xint#1{\mathchoice
+	{\XXint\displaystyle\textstyle{#1}}%
+	{\XXint\textstyle\scriptstyle{#1}}%
+	{\XXint\scriptstyle\scriptscriptstyle{#1}}%
+	{\XXint\scriptscriptstyle\scriptscriptstyle{#1}}%
+	\!\int}
+\def\XXint#1#2#3{{\setbox0=\hbox{$#1{#2#3}{\int}$}
+		\vcenter{\hbox{$#2#3$}}\kern-.5\wd0}}
+\def\ddashint{\Xint=}
+\def\dashint{\Xint-}
+
 %%% Misc
@@
 \begin{document}
-\title{Minimax-optimal regression over Sobolev Spaces with Neighborhood Graphs}
+\title{Note on: Avoiding Pitfalls when Thinking about Sobolev Spaces}
 \author{Alden Green}
@@
 
-\section{Introduction}
-\label{sec:introduction}
+The purpose of this note is to record some errors we've made when thinking about Sobolev spaces, and continuum estimators defined using their norms (thin-plate splines).
+The goal is to avoid repeating these errors in the future.
 
-\textbf{(1) Geometric graphs.} Suppose one observes data $X_1,\ldots,X_n$ sampled independently from an unknown distribution $P$, and forms a geometric graph $G$---with edges $e_{ij}$ corresponding to proximity between samples $X_i$ and $X_j$---over the observed data.
-Geometric graphs encode information about $P$ in an extremely general manner, and have thus been leveraged to conduct many different fundamental statistical tasks.
-These include clustering, semi-supervised learning, classification, and regression---both estimation and goodness-of-fit testing.
-Though much theoretical work has been done on the consistency of graph-based learning methods---and more recently, rates of convergence have been established for some problems---little is known so far about their optimality, even for classic statistical tasks.
+\begin{itemize}
+	\item \textbf{Lesson 1: Pick your representative.} Elements $u$ of Sobolev spaces $W^{k,p}(\Omega)$ are equivalence classes.
+Functions $f$ and $g$ are equivalent if they disagree only a set of Lebesgue measure-zero.
+In order to speak of pointwise evaluation of some $u \in W^{k,p}(\Omega)$, we must pick first pick a representative $f^{\star} \in u$.
+	\begin{itemize}
+		\item When $pk > d$, by the Sobolev Embedding Theorem every $u \in W^{k,p}(\Omega)$ has a continuous representative, and the convention is to take $f^{\star}$ to be that continuous representative.
+		\item When $pk \leq d$, $W^{k,p}(\Omega)$ no longer embeds into $C(\Omega)$. In this case, we choose $f^{\star}$ to be the \emph{precise representative} of $u$, defined for $x \in \Omega$ as
+		\begin{equation*}
+		f^{\star}(x) :=
+		\begin{dcases}
+		\lim_{r \to 0} \biggl\{\frac{1}{\nu(B(x,r))} \int_{B(x,r)} u(x) \,dx\biggr\},& ~~\textrm{if the limit exists} \\
+		0,& ~~\textrm{otherwise.}
+		\end{dcases}
+		\end{equation*}
+		Note that by the Lebesgue Differentation Theorem, the set $E$ of Lebesgue points of any $u \in W^{k,p}(\Omega) \subset L^1(\Omega)$ satisfies $\nu(\Omega\setminus E) = 0$.
+Thus $f^{\star}$ is indeed a representative of $u$.
+Moreover when $2k > d$, $f^{\star}$ is continuous.
+	\end{itemize}
+	Once we have picked representatives, we can make sense of the objective function
+	\begin{equation*}
+	R(u) = \|Y - u\|_n^2 + \lambda |u|_{W^{k,p}(\Omega)}^2.
+	\end{equation*}
+	Henceforth when we say ``function $f \in W^{k,p}(\Omega)$'', we are referring to the precise representative $f^{\star}$ of an equivalence class $u$.
+	\item \textbf{Lesson 2: Be precise about why the smoothing spline problem is well-posed when $2k > d$.} Let $p = 2$.
+When $2k > d$, the Hilbert-Sobolev space $H^k(\Omega) = W^{k,2}(\Omega)$ can be continuously embedded into $C^{j,\gamma}(\Omega)$ with $\gamma = \floor{d/2} + 1 - d/2$ and $j = k - \gamma - d/2$ (Here for convenience we have assumed $d$ is odd.) For the moment set $d = 1$ and $\Omega = [0,1]$.
+When $d = 1$, observe that $\gamma = 1/2$ and $j = k - 1$.
+Therefore any function $f \in H^k(\Omega)$ is absolutely continuous, with continuous derivatives $f',f'',\ldots,f^{(k- 1)}$, and with $k$th weak derivative bounded in $L^2([0,1])$ norm.
+These continuity properties ensure that the \emph{smoothing spline} optimization problem,
+	\begin{equation}
+	\label{eqn:smoothing_spline}
+	\min_{H^k([0,1])} \|Y - u\|_n^2 + \lambda |u|_{W^{k,p}(\Omega)}^2,
+	\end{equation}
+	is well-defined.
+By this we mean that there exists a unique function $\wt{u} \in H^k([0,1])$ which achieves the minimum in~\eqref{eqn:smoothing_spline}.
+Much more is known about the properties of $\wt{u}$, but that is not important for this discussion.
+	\item \textbf{Lesson 3: Be precise about why the smoothing spline problem is ill-posed when $2k \leq d$.} Let's return to the general multivariate setting, where $d$ is arbitrary and $\Omega$ is a regular domain.
+Suppose $2k \leq d$.
+Now we know problem~\eqref{eqn:smoothing_spline} is ill-posed.
+My previous explanation of this facet would have been ``it is ill-posed because you can put spikes at the data'' but that is not the case--we have already ruled out changes on a set of measure zero by picking a precise representative.
+The correct explanation is that $R(u)$ is not a continuous function of $u$ with respect to the topology induced by $H^k(\Omega)$ norm.
+For instance, take a sequence of bump function $g_h = \textcolor{red}{(?)}$.
+Clearly $g_h(x) = 0$ for any $x \in \Omega\setminus B(0,h)$.
+On the one hand, a standard calculation will show that $g_h \to 0$ as $h \to \infty$, where crucially convergence is meant in the Sobolev norm.
+On the other hand, $g_h(0) = 1$ for all $h$, and so $R(g_h) \not\to R(0)$.
+For this reason~\eqref{eqn:smoothing_spline} does not have a minimizer in $H^k(\Omega)$.
 
-\textbf{(2) What we study: Regression and Laplacian eigenmaps.} In this paper we focus on regression, where in addition to the design points $X_1,\ldots,X_n$ one observes real-valued responses $Y_1,\ldots,Y_n$, and seeks to infer the regression function $f_0(x) := E[Y|X = x]$.
-The method we will study is \emph{Laplacian eigenmaps}, first introduced by \textcolor{red}{(Belkin 2003)}, which constructs an estimator $\wh{f}$ using eigenvectors of the graph Laplacian $L$ as features in a least-squares problem.
-The graph Laplacian $L$ is a difference operator, acting on functions $f: X \to \Reals$,
-\begin{equation}
-\label{eqn:graph_laplacian}
-Lf(X_i) = \sum_{j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)e_{ij},
-\end{equation}
-The Laplacian $L$ can be seen as a discretization of the Laplace-Beltrami operator \textcolor{red}{(...)}.
-The eigenvectors of $L$ are an orthonormal basis of $L^2(X)$, which can seen as estimates of the $L^2(P)$ orthonormal basis \textcolor{red}{(...)}.
-Their corresponding eigenvalues  are estimates of~\textcolor{red}{(...)}, and provide a notion of smoothness for each eigenvector---the smaller the eigenvalue, the smoother the eigenvector.
-
-\textbf{(3) Why we study.} Laplacian eigenmaps is thus a type of \emph{spectral series} estimator.
-Spectral series estimators are one of the most classical ways to perform non-parametric regression.
-Traditional spectral series estimators use an \emph{a priori defined basis} \textcolor{red}{(...)}, and are typically defined with respect to some reference measure $Q$.
-In contrast, eigenvectors of the graph Laplacian are data-dependent, and adapt to the geometry of the distribution $P$ in a rich manner.
-
-The study of (classical) spectral series estimators is rich, and their theoretical optimality is well-understood \textcolor{red}{(references)}.
-However, in practice these estimators suffer from some serious drawbacks.
-
-\textbf{(4) Our contributions.} Laplacian eigenmaps avoid these drawbacks \textcolor{red}{(...)} On the other hand their statistical properties are not as well understood \textcolor{red}{(...)}.
-The primary contribution of our paper is to fill this theoretical gap, by answering the following question:
-
-\begin{quote}
-	\textcolor{red}{(TODO)}
-\end{quote}
-
-We show that under various classical assumptions on the regression function $f_0$ and design distribution $P$, Laplacian eigenmaps is statistically minimax optimal.
-\textcolor{red}{(TODO): Add a richer summary of your major results here.}
-
-\textbf{(5) Organization.}
-
-\section{Setup and Background}
-\label{sec:setup_main_results}
-
-In this section, we give a precise setup of the non-parametric regression problems and methods we will subsequently consider, review the minimax rates for non-parametric regression over Sobolev spaces, and summarize our main results.
-
-\subsection{Non-parametric regression with Laplacian Eigenmaps}
-\label{sec:regression_laplacian_eigenmaps}
-
-\textbf{(1) Formal setup.}
-Throughout, we assume that we observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_N$ are drawn i.i.d from a distribution $P$ supported on a compact set $\mc{X} \subseteq \Rd$, and
-\begin{equation*}
-Y_i = f_0(X_i) + \varepsilon_i,
-\end{equation*}
-with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise with unit variance.
-
-\textbf{(2) Laplacian eigenmaps.}
-
-\subsection{Sobolev Classes}
-\label{sec:sobolev}
-We now review the Sobolev classes $H^s(\mc{X})$.
-
-\textbf{(1) Sobolev norms, semi-norms, and balls.}
-
-\textbf{(2) Zero-trace conditions.}
-
-\subsection{Minimax Rates}
-\label{subsec:minimax_rates_sobolev}
-
-\textbf{(1) Estimation rates.}
-
-\textbf{(2) Testing rates.}
-
-\textbf{(3) Conditions under which spectral series methods achieve these rates.}
-
-\textbf{(4) Comparison with the conditions necessary for Laplacian Eigenmaps.}
-
-\subsection{Our contributions}
-The following items summarize our major results.
-\begin{itemize}
-	\item \textbf{Minimax optimal estimation.} With high probability, $\norm{\wh{f} - f_0}_{n}^2 \lesssim n^{-2/(2 + d)}$ over all $f_0 \in H^1(\Xset;M)$. 
-	\item \textbf{Minimax optimal testing.}
-	A test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in H^1(\Xset;M)$ satisfies $\norm{f_0}_{\Leb^2(\Xset)} \gtrsim n^{-2/(4 + d)}$ and $d < 4$.
-	\item \textbf{Taking advantage of higher-order smoothness conditions.} When the regression function $f_0 \in H^s(\Xset)$ and additionally the density $p \in C^{s - 1}(\mc{X})$, with high probability the error $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2s/(2s + d)}$ over all $f_0 \in H_0^s(\Xset;M)$.
-Additionally, a test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in H_0^s(\Xset;M)$ satisfies $\norm{f_0}_{\Leb^2(\Xset)} \gtrsim n^{-2s/(4s + d)}$ and $d < 4$.
-	\item \textbf{Manifold adaptivity.}
-	If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$, each of the aforementioned rates hold with $d$ replaced by $m$.
-\end{itemize}
-\textbf{(1) Major takeaway.} Viewed as a whole, our results show that Laplacian eigenmaps retain the strong theoretical properties of classic orthogonal series approaches, while also automatically \emph{adapting} to an unknown design distribution $P$.
-
-\subsection{Related Work}
-\begin{itemize}
-	\item \cite{lee2016} considers a similar estimator --- the notable difference being the normalization of the Laplacian $L$ --- but get suboptimal rates for the supervised problem. 
-	\item \cite{zhou2011} consider a similar estimator, but in the semi-supervised setting.
-They consider the limit as the number of unlabeled points grows to infinity, in which case the estimator is simply a classic spectral series estimator, and the analysis is standard.
+	NB: This is a particular example of a general mistake I often make, which is forgetting that convergence of a sequence in Sobolev norm does not imply pointwise convergence, either of the function $f$ (when $2k \leq d$) or of the $j$th derivative of the function (when $2(k - j) \leq d$.).
+	\item \textbf{Lesson 4: Practical implementation.} In our AISTATS paper,  I make the following comment:
+	\begin{quotation}
+		...any practical implementation of~\eqref{eqn:smoothing_spline} will be inconsistent [when $d \geq 2$ and $k = 1$.]
+	\end{quotation}
+	This comment is too imprecise to be interpretable.
+Let's give it some meaning.
+Numerically, one could try to solve~\eqref{eqn:smoothing_spline} by discretizing the space $\Omega$ and solving a discrete version of~\eqref{eqn:smoothing_spline}.
+\textcolor{red}{This discrete problem} will always be well-posed, meaning it will always admit a unique solution.
+Unfortunately, if one takes the discretization of $\Omega$ to be \textcolor{red}{sufficiently fine} (compared to the number of samples $n$) the resulting estimator $\check{u}$ will be an inconsistent estimator of $f_0$.
+	\item \textbf{Lesson 5: Be more imaginative.} I tend to think of Sobolev spaces as containing continuous and continuously differentiable functions, up to local deviations (spikes).
+This intuition is only appropriate when $2k > d$.
+For instance, the fractional Sobolev space $H^{1/2}([0,1])$ contains step-functions.
+When $d \geq 2$, weak differentiability does not imply continuity.
 \end{itemize}
 
-\section{Minimax Optimality of Laplacian Eigenmaps}
-\label{sec:minimax_optimal_laplacian_eigenmaps}
-
-In this section, we will show that the estimator $\wh{f}$, and a test using the statistic $\wh{T}$, achieve optimal estimation and goodness-of-fit testing rates over Sobolev classes.
-Throughout this section, we will assume $\mc{X}$ is a \textcolor{red}{regular, $d$-dimensional domain} and $P$ has a \textcolor{red}{regular density with respect to Lebesgue measure.}
-
-\subsection{First-order Sobolev classes}
-\label{sec:first_order_sobolev_classes}
-To begin, we handle the case where $f_0 \in H^1(\mc{X})$.
-
-\textbf{(1) Laplacian eigenmaps is an optimal estimator.} Under appropriate choices of $r$ and $\kappa$, the estimator $\wh{f}$ has (in-sample) optimal risk (up to constant factors) $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2/(2 + d)}$.
-
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_estimation_fo}
-	\textcolor{red}{(TODO)}
-	\begin{equation}
-	\label{eqn:laplacian_eigenmaps_estimation_fo}
-	\|\wh{f} - f_0\|_n^2 \lesssim n^{-2/(2 + d)}
-	\end{equation}
-\end{theorem}
-Some remarks:
+\textcolor{red}{(TODO):}
 \begin{itemize}
-	\item \textbf{(1.1) How to tune hyper-parameters.} In practice, one typically tunes the hyper-parameters of an algorithm by sample-splitting or (more usually) by cross-validation.
-However, because $\wh{f}$ is defined only in-sample such approaches cannot be straightforwardly applied to select the graph radius $r$, or number of eigenvectors $\kappa$.
-We discuss out-of-sample extensions of $\wh{f}$ later, in Section~\ref{sec:out_of_sample}.
-	\item \textbf{(1.2) The dependence on failure probability can be improved.} The upper bound on in-sample risk given by Equation~\eqref{eqn:laplacian_eigenmaps_estimation_fo} holds with probability $1 - C\delta$, and depends on a factor of $1/\delta$.
-When $f \in C^1(\mc{X})$, we can show that this depends on a factor of $(1 + 1/\delta)$, and holds with probability at least $1 - C\delta^2/n$.
+	\item Note that when $2k > d$, and $d > 1$, the $(k-1)$st partial derivatives of $f \in H^k(\Omega)$ may not be classically defined, and yet~\eqref{eqn:smoothing_spline} is still well-defined.
+Find what continuity properties suffice for~\eqref{eqn:smoothing_spline} to be well-defined.
+	\item Give an example of a \textcolor{red}{reasonable function} $f \in H^1([-1,1]^2)$ which is discontinuous. A brief Google search finds examples which seem straight out of the mind of Georg Cantor.
 \end{itemize}
 
-\textbf{(2) Laplacian eigenmaps is an optimal test.} Consider the test $\phi = \1\{\wh{T} \geq t\}$, where $t$ is the data-dependent threshold~\textcolor{red}{(...)}.
-The choice of $t$ guarantees that $\phi$ is a level-$\alpha$ test.
-Under appropriate choices of $r$ and $\kappa$, the test $\phi$ has non-negligible power (up to constant factors) against alternatives separated the null by at least $\|f_0\|_{\Leb^2(\mc{X})} \gtrsim n^{-4/(4 + d)}$.
-
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_testing_fo}
-	\textcolor{red}{(TODO)}
-\end{theorem}
-Some remarks:
-\begin{itemize}
-	\item \textbf{(2.1) When $d \geq 4$, the testing problem is different.} When $d \geq 4$, it is not the case that $\Leb^4(\mc{X})$ continuously embeds into $H^1(\mc{X})$, and testing rates over $H^1(\mc{X})$ are not known.
-On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then simply taking $T = \|Y\|_n^2$ achieves the optimal rates.
-	\item \textbf{(2.2) Both of the remarks after Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} also apply to testing.}
-\end{itemize}
-
-\subsection{Higher-order Sobolev classes}
-\label{sec:higher_order_sobolev_classes}
-In this section, we handle the case where $f_0 \in H_0^s(\mc{X})$.
-
-\textbf{(1) Laplacian eigenmaps is an optimal estimator.}  Under appropriate choices of $r$ and $\kappa$, the estimator $\wh{f}$ has (in-sample) optimal risk (up to constant factors) $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2s/(2s + d)}$.
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_estimation_ho}
-	Under appropriate conditions, with probability at least $1 - \delta$,
-	\begin{equation}
-	\label{eqn:laplacian_eigenmaps_estimation_ho}
-	\|\wh{f} - f_0\|_n^2 \lesssim n^{-2s/(2s + d)}
-	\end{equation}
-\end{theorem}
-Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that Laplacian eigenmaps is a rate-optimal estimator over Sobolev classes, for all scalings of $s$ and $d$.
-Some other remarks:
-\begin{itemize}
-	\item \textbf{(1.1) Optimal rates, no RKHS.} It is worth pointing out that we do not require that $2s > d$, a condition often seen in the literature.
-When $2s \leq d$, the Sobolev space $H^s(\mc{X})$ is quite irregular.
-For instance, it is not a Reproducing Kernel Hilbert Space, nor does it embed into any H\"{o}lder space.
-This renders certain versions of the nonparametric regression problem ill-posed---for instance when the design points $X$ are assumed to be fixed, or loss is measured in $\Leb^{\infty}$ norm.
-It also takes many estimators ``off the table'', most notably including RKHS-based methods such as smoothing splines.
-However, at these scales traditional spectral series estimators still obtain the optimal rates for random design regression measured in $\Leb^2$ norm.
-Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} shows that the same is true with respect to Laplacian eigenmaps.
-	\item \textbf{(1.2) The zero-trace condition.} The zero-trace condition can likely be weakened.
-However, requiring some boundary condition on $f_0$ and its first through $(k -1 )$th-order derivatives is unavoidable.
-\textcolor{red}{Explain why: graphs are implicitly boundaryless objects, and the eigenvectors of $L$ converge to those eigenvectors of $\Delta_P$ which satisfy boundary conditions.
-Repeat that this is a standard assumption made when analyzing spectral series estimators}.
-	\item \textbf{(1.3) The bounded norm and smooth density conditions.} We require that $f_0$ be bounded in the Sobolev norm $\|f_0\|_{H^s(\mc{X})}$---as opposed to the semi-norm $|f_0|_{H^s(\mc{X})}$--- and that $p \in C^{s - 1}(\mc{X})$.
-Both requirements are related and fundamental, as we shall now explain.
-As previously mentioned, $L$ converges to a weighted-version of the Laplace operator $\Delta_P$ (also known as a Fokker-Planck) operator.
-The operator $\Delta_P$ includes both first and second-order derivatives,
-	\begin{equation}
-	\label{eqn:fokker_planck_1}
-	\Delta_Pf= \frac{-1}{p} \mathrm{div}(p \nabla f) = \Delta f - \frac{\nabla p^{\top} \nabla f}{p},
-	\end{equation}
-	and therefore a function $f \in H_0^s(\mc{X})$ has finite energy $\|\Delta_P f\|_{P}^2$ only if the first-order partial derivatives of $f$ and $p$ are bounded.
-\textcolor{red}{This discussion---and in particular the last sentence---are obviously not quite correct.
-This is in part because you realize now that the requirements are not in fact fundamental.
-Instead, the requirement should merely be that $\|\Delta_P^kf\|_{\Leb^2(P)}^2$ is bounded---here $k = s/2$ and we assume $s$ is even---and the upper bounds should be in terms of this semi-norm.}.
-	\item \textbf{(1.4) Both of the remarks after Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} continue to apply.}
-\end{itemize}
-
-\textbf{(1.4) Rates for testing with higher-order smoothness assumptions.} For testing, we can get the optimal rate $n^{-4s/(4s + d)}$ when $d \leq 4$.
-When $4 < d \leq 4s$, we get the rate~\textcolor{red}{(...)} Recall that the rate $n^{-4s/(4s + d)}$ is optimal whenever $d \leq 4s$.
-Therefore there are scalings of $s$ and $d$ for which we do not get the optimal rate.
-We do not know whether this due to looseness in our proof techniques, or a problem with the test itself.
-
-\subsection{Analysis}
-\label{sec:analysis}
-
-\textbf{(1) Bias and variance terms, for estimation and testing}.
-We analyze the estimation error of $\wh{f}$ by first condition on the design points $X$ and computing \emph{design-dependent} bias and variance terms.
-For estimation, we have
-\begin{equation*}
-\Ebb\bigl[\|\wh{f} - f_0\|_n^2|X\bigr] \leq C\frac{\dotp{L^s f_0}{f_0}_n}{\lambda_{\kappa}^s} + \frac{\kappa}{n}
-\end{equation*}
-For testing, we have~\textcolor{red}{(...)}.
-
-In the above, the semi-norm $\dotp{L^s f_0}{f_0}_n$ and eigenvalue $\lambda_{\kappa}$ are random variables that depend the design $X$.
-It remains to derive suitable upper and lower bounds on these quantities.
-
-\textbf{(2) Upper bound on the semi-norm. } Lemma~\ref{lem:graph_seminorm} gives a sufficient upper bound on $\dotp{L^s f_0}{f_0}_n$. Take $\beta := s - 1$. 
-\begin{lemma}
-	\label{lem:graph_seminorm} 
-	Suppose $f \in H_0^s(\mc{X})$, that $p \in C^{\beta}(\mc{X})$, and that $r \gtrsim n^{-1/d}$. Then with probability at least $1 - \delta$, it holds that 
-	\begin{equation}
-	\label{eqn:graph_seminorm_1}
-	\dotp{L^s f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^s} \|p\|_{C^{\beta}} + \textcolor{red}{(?)}
-	\end{equation}
-	Therefore, if additionally $r \gtrsim n^{-1/(2\beta + d)}$, then 
-	\begin{equation}
-	\label{eqn:graph_seminorm_2}
-	\dotp{L^s f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^s} \|p\|_{C^{\beta}}
-	\end{equation}
-\end{lemma}
-
-\textbf{(3) Lower bound on the eigenvalue.} On the other hand, recent work \textcolor{red}{(Burago14, Shi16, Garcia Trillos 18)} has analyzed the convergence of $\lambda_{k}$ towards $\lambda_{k}(\Delta_P)$.
-They provide explicit bounds on the relative error $|\lambda_{k} - \lambda_{k}(\Delta_P)|/\lambda_{k}(\Delta_P)$, showing that the relative error is small for sufficiently large $n$, small $\varepsilon$, and any $1 \leq k \leq n$ that is not too large relative to $n$.
-These results are actually stronger than are necessary to establish Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_estimation_ho}---note that we need only show that $\lambda_{\kappa}$ is lower bounded by the right order of magnitude---but unfortunately they all assume $P$ is supported on a manifold without boundary.
-We will instead use the results of \textcolor{red}{(Green 2021)}, whose assumptions match our own, and who give a weaker bound on $\lambda_k/\lambda_k(\Delta_P)$ that will nevertheless suffice for our purposes.
-
-\begin{lemma}
-	\textcolor{red}{(TODO)}
-	With probability at least $1 - \textcolor{red}{(?)}$,
-	\begin{equation*}
-	\lambda_k \geq c \cdot \min\Bigl\{\lambda_k(\Delta_P), \frac{1}{r^{2}} \Bigr\}
-	\end{equation*}
-\end{lemma}
-By Weyl's Law, $c k^{2/d} \leq \lambda_{k}(\Delta_P) \leq Ck^{2/d}$, and as a result $\lambda_{\kappa} \geq C\lambda_{\kappa}(\Delta_P)$ as long as $r \geq \kappa^{-1/d}$.
-
-\textbf{(1.2) Traps we avoid.} It is tempting to use recent results regarding spectral convergence to analyze the estimator $\wh{f}$ and test $\wh{T}$.
-This suffers from an accumulation of error problem.
-
-It is also key that we directly analyze the semi-norm $\dotp{L^s f}{f}_n$, rather than invoking the pointwise convergence of $L^{s}f \to \Delta_P^{s}f$ to obtain a bound on $\dotp{L^s f}{f}_n$.
-Recall that we assume $f$ has bounded derivatives up to (and including) order $s$.
-This suffices to gain control of the semi-norm $\dotp{L^s f}{f}_n$, because $\dotp{L^s f}{f}_n$ approximates the $\Leb^2$ norm of an order $s$ differential operator, and so we have exactly as many derivatives as we need.
-On the other hand, $L^s$ approximates an order $2s$ differential operator, and $\Delta_P^sf$ is only a well-defined limiting object if we assume $f$ has order $2s$ derivatives.
-
-\section{Manifold Adaptivity}
-\label{sec:manifold_adaptivity}
-
-The optimal rates for nonparametric regression suffer from a bona fide curse of dimensionality, and are thus difficult to reconcile with the practical success of many methods for nonparametric regression.
-One explanation for this seeming paradox is the \emph{manifold hypothesis}, where the design points $X$ are assumed to lie on or near a manifold $\mc{X}$ of intrinsic dimension $m \ll d$.
-Under the manifold hypothesis nonparametric regression is demonstrably easier, and in particular it is known \citep{bickel2007,ariascastro2018} that in this setting the optimal rates over $H^s(\Xset)$ scale like $n^{-2s/(2s + m)}$ (for estimation) and $n^{-4s/(4s + m)}$ (for testing).
-
-On the other hand, a theory has been developed~\citep{belkin03,belkin05,niyogi2013} establishing the the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
-In this section, we contribute to this line of work by showing that under the manifold hypothesis, Laplacian eigenmaps achieve the sharper minimax estimation and testing rates.
-First, we give some relevant background on Riemmanian manifolds, and Sobolev classes defined upon them.
-
-\subsection{Riemmanian manifolds}
-
-\subsection{Error rates under the manifold hypothesis}
-
-\section{Out-of-sample error}
-\label{sec:out_of_sample}
-Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity} highlights the statistical optimality of nonparametric regression with Laplacian eigenmaps.
-However, with regards to estimation, the error has always been measured in $\|\cdot\|_n$ norm---in words, the average squared loss incurred over the $n$ observed design points.
-This is somewhat unusual.
-Instead, it is more typical to consider $L^2(P)$ error---meaning error with respect to the $\|\cdot\|_{P}$ norm---in large part because $L^2(P)$ error is closely tied to the expected prediction error an estimator would make at a new point $X_{n + 1}$.
-
-Of course, the Laplacian eigenmaps estimator is only defined at the observed design points $X_1,\ldots,X_n$, and to measure its error in $L^2(P)$ norm we must first give a means of extending it to the rest of $\Xset$.
-We propose a simple method, essentially kernel smoothing, to do the job, which can be applied to any estimator at data, including Laplacian eigenmaps.
-We show that the smoothed version of $\wh{f}$ has optimal $L^2(P)$ error.
-
-\subsection{Extending Laplacian Eigenmaps with Kernel Smoothing}
-
-\begin{theorem}
-	\label{thm:kernel_smoothing}
-	\textcolor{red}{(TODO)}
-\end{theorem}
-
-Some remarks:
-\begin{itemize}
-	\item \textbf{Tuning parameters by sample-splitting.}
-	\item \textbf{Other extensions.} Comment that there does not exist any canonical way to extend Laplacian based estimators such as $\wh{\theta}$ to an ambient domain.
-Point out that there exist various proposals to extend eigenvectors in the literature, such as $1$-nearest-neighbor regression (piecewise constant extrapolation over Voronoi cells), convolution with a specialized kernel, or Nystr\"{o}m extension.
-Explain that each of these is designed for a highly specialized purpose, such as suitability for mathematical analysis or computational efficiency, and in any case are suggested only for eigenvectors rather than a regression estimate.
-State explicitly that we choose kernel smoothing for its simplicity, ubiquity, and strong theoretical properties under our assumptions, but that considerations of extensions for specifically tailored to Laplacian based estimators could be of interest.
-\end{itemize}
-
-\subsection{Why not just do kernel smoothing?}
-On the other hand, the allusion to kernel smoothing raises a very natural question: why should we prefer to first compute $\wh{\theta}$ and smooth it out, as opposed to simply performing kernel smoothing over the original responses $Y$.
-We now answer this question, by \textcolor{red}{(...)}.
-
-\section{Experiments}
-\label{sec:experiments}
-
-\section{Discussion}
-\label{sec:discussion}
-
-\bibliographystyle{plainnat}
-\bibliography{../../../graph_testing_bibliography}
-
 \end{document}

### 2021-02-24 notes/misc/insights_and_connections.tex
Added a note on some pitfalls to avoid when thinking about Sobolev spaces.

@@
 	\end{itemize}
-	\item We should also give some insight into the structure of $V_N$.
+	\item \textcolor{red}{(TODO)}: We should also give some insight into the structure of $V_N$.
 For instance, what are the smoothness properties of functions $u \in V_N$?
@@
 How do these depend on the density $p$?
+	\item \textcolor{red}{(TODO)}: Some grumpy advisor might say that we need to justify why anyone should care about relating $\wh{u}$ to $\check{u}$.
+One way to do this is by arguing that methodologically~\eqref{eqn:bandlimited_laplacian_smoothing} is superior to~\eqref{eqn:laplacian_smoothing}, for instance because it is well defined over all of $\Omega$ rather than only $\Omega_N$.
+Another way to do this is by arguing that the representation~\eqref{eqn:bandlimited_laplacian_smoothing} buys us something theoretically.
+Three example things it might buy are:
+	\begin{enumerate}[(1)]
+		\item It may be possible to relate~\eqref{eqn:bandlimited_laplacian_smoothing} to an equivalent kernel,
+		\item It is far easier to analyze~\eqref{eqn:bandlimited_laplacian_smoothing} than directly analyzing Laplacian smoothing.
+Hopefully, we can easily get sharp upper bounds for more interesting situations than we have analyzed thus far.
+For instance, when $p$ or $\Omega$ are complicated, instead of being a smooth function with a smooth boundary, respectively.
+		\item It should be easy to prove a lower bound on the error of~\eqref{eqn:bandlimited_laplacian_smoothing} when $d > 4$. 
+	\end{enumerate}
 \end{itemize}

### 2021-02-24 code/plot.R
Updated plots.

@@ -1,8 +1,9 @@
+library(dplyr)
 library(ggplot2)
 library(gridExtra)
 
 ### Plot. ###
-save_directory <- "data/AISTATS/cosine/d2_largesample" # Please change this to whichever directory you prefer.
+save_directory <- "data/AISTATS/cosine/d1_largesample" # Please change this to whichever directory you prefer.
 plot_directory <- file.path(save_directory,"plots")
 dir.create(plot_directory)
 
@@ -95,8 +96,11 @@ for(jj in 1:length(methods))
   {
     mse_ii_jj <- mse[[ii]][[jj]]
     best_mse[ii] <- min(rowMeans(mse_ii_jj))
-    minimax_mse[ii] <- 4*ns[ii]^{-2/(2 + d)}
+    minimax_mse[ii] <- ns[ii]^{-2/(2 + d)}
   }
+  
+  # Rescale minimax mse to match intercept with best_mse
+  minimax_mse <- minimax_mse * (best_mse[1]/minimax_mse[1])
   plot_dfs_best_mse[[jj]] <- data.frame(x = ns, y = best_mse, z = minimax_mse) 
   
   # fitted slope
@@ -126,7 +130,7 @@ ltys <- c(2,3)
 
 # shadow plot
 plot(x = ns, xlim = xlims, ylim = ylims,
-     log = "xy", xlab = "n", ylab = "mse", main = title, cex.main = 2, cex.lab = 2, cex.axis = 1.5)
+     log = "xy", xlab = "Sample size", ylab = "Mean squared error", main = title, cex.main = 2, cex.lab = 2, cex.axis = 1.5)
 
 # add points and lines
 for(jj in 1:length(methods))

### 2021-02-24 manuscripts/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
Just use one CMU label
Halfway through responding to Ryan's changes.

@@
 Laplacian smoothing \citep{smola2003} is a penalized least squares estimator, defined over a graph.
-Let $G = (V,W)$ be a weighted undirected graph with vertices $V=\{1,\ldots,n\}$, associated with $\{X_1,\ldots,X_n\}$, the Laplacian smoothing estimator \smash{$\wh{f}$} is given by
+Letting $G = (V,W)$ be a weighted undirected graph with vertices $V=\{1,\ldots,n\}$, associated with $\{X_1,\ldots,X_n\}$, the Laplacian smoothing estimator \smash{$\wh{f}$} is given by
 \begin{equation}
@@
 It's fine, I wouldn't change anything because it would involve explaining a lot more, I'm just pointing this out.
-(Throughout we assume $M \geq n^{-1/2}$, as otherwise the minimax rate is just the parametric rate $n^{-1}$.)
+(Throughout we assume $M \geq n^{-1/2}$---as otherwise the minimax rate is just the parametric rate $n^{-1}$.)
 
@@
 	\label{thm:laplacian_smoothing_estimation1}
-  Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ has dimension $d < 4$.
+  Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ has dimension $d < 4$ and where $M \leq n^{1/d}$.
 Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume the neighborhood graph \smash{$G_{n,r}$} is computed with a kernel $K$ satifsying \ref{asmp:kernel}.
 There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_estimation}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$}, satisfies
-	\begin{equation*}
+	\begin{equation}
+	\label{eqn:laplacian_smoothing_estimation1}
 	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(2 + d)} n^{-2/(2 + d)},
-	\end{equation*}
-  with probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c \min\{M^{d/(2d + 4)} n^{d/(2+d)},n\})$.
+	\end{equation}
+  with probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c (M^2n)^{d/(2 + d)})$.
 % RJT: Alden, please verify that what I wrote is correct (I reorganized the way we explained this, added for $n \geq N$ ...).
+% AJG: Verified.
 \end{theorem}
@@
 % RJT: Alden, please add a section to the supplement to explain this part.
+% AJG: Added.
+	\item The assumption $M \leq n^{1/d}$ guarantees that the bound in~\eqref{eqn:laplacian_smoothing_estimation1} is meaningful, i.e.
+that the right hand side of~\eqref{eqn:laplacian_smoothing_estimation1} is of at most a constant order.
 	\item The lower bound on $r$ imposed in condition \ref{asmp:ls_kernel_radius_estimation} is compatible with practice, where by far the most common choice of radius is the connectivity threshold \smash{$r \asymp (\log(n)/n)^{1/d}$}, which makes \smash{$G_{n,r}$} as sparse as possible while still being connected, for maximum computational efficiency.
@@
 	\label{thm:laplacian_smoothing_estimation2}
-  Under the assumptions of Theorem~\ref{thm:laplacian_smoothing_estimation1}, if $\Xset$ has dimension $d = 4$ and \smash{$r \asymp (\log n/n)^{1/4}$}, then we obtain 
+  Under the assumptions of Theorem~\ref{thm:laplacian_smoothing_estimation1}, if $\Xset$ has dimension $d = 4$, \smash{$r \asymp (\log n/n)^{1/4}$} and \smash{$\rho = M^{-2/3}(nr^6)^{-1}(\log n/n)^{1/3}$}, then we obtain
 	\begin{equation*}
@@
 	\end{equation*}
-with the same probability guarantee as in Theorem~\ref{thm:laplacian_smoothing_estimation1}. If the dimension $\Xset$ is $d \geq 5$, TODO
+with the same probability guarantee as in Theorem~\ref{thm:laplacian_smoothing_estimation1}.
+If the dimension of $\Xset$ is $d \geq 5$, \smash{$r \asymp (\log n/n)^{1/d}$} and \smash{$\rho =  M^{-2/3}(nr^{2 + d})^{-1}n^{-4/(3d)}$}, then
+	\begin{equation*}
+	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{4/3} \biggl(\frac{\log n}{n}\biggr)^{-4/(3d)},
+	\end{equation*}
 % RJT: Alden, please verify what I wrote is correct, and fill in the result for d \geq 5.
+% AJG: Verified and filled in.
 \end{theorem}
-
 This mirrors the conclusions of \citet{sadhanala16} who investigate estimation rates of Laplacian smoothing over the $d$-dimensional grid graph.
@@
 Oh and if you're already going to redo one of these plots, you can satisfy my unreasonably nitpicky brain and change the x- and y-axes on all of them to use words rather than symbols/acronyms: "Sample size" and "Mean squared error".
+% AJG: Verified, and redone with requested changes.
 \end{figure*}
@@
 The next theorem shows that this test also has small Type II error, uniformly over all $f_0$ that are separated from $0$ by at least the critical radius given in \eqref{eqn:sobolev_space_testing_critical_radius}.
-For this to hold, we will require a tighter range of scalings for the radius $r$.
+For this to hold, we will require a tighter range of scalings for the graph radius $r$.
 \begin{enumerate}[label=(R\arabic*)]
@@
 Discretization is not the only way to make the problem \eqref{eqn:thin_plate_spline} more tractable: for instance, one can replace the penalty \smash{$\int_{\Xset} \|\nabla f(x)\|_2^2 \,dx$} with a stricter choice like \smash{$\esssup_{x \in \Xset} \|\nabla f(x)\|_2$}, or conduct the optimization over some finite-dimensional linear subspace of $H^1(\Xset)$ (i.e., use a sieve).
-While these solutions do improve the statistical properties of \smash{$\wt{f}$} for $d > 1$ (see e.g., \citet{birge1993,birge1998,vandergeer2000}), Laplacian smoothing is generally speaking much simpler and more computationally friendly, and in addition, the other approaches are specifically tailored to the domain $\Xset$, in stark to \smash{$\wh{f}$}.
+While these solutions do improve the statistical properties of \smash{$\wt{f}$} for $d > 1$ (see e.g., \citet{birge1993,birge1998,vandergeer2000}), Laplacian smoothing is generally speaking much simpler and more computationally friendly.
+In addition, the other approaches are specifically tailored to the domain $\Xset$, in stark contrast to \smash{$\wh{f}$}.
 

### 2021-02-24 manuscripts/laplacian_smoothing/AISTATS/supplement_minimax_optimal_laplacian_smoothing.tex
Halfway through responding to Ryan's changes.

@@
 Our main theorems (Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing_manifold}) all follow the same general proof strategy, centered on conditioning.
-In Section~\ref{sec:fixed_graph_error_bounds}, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to (functionals of) the graph $G$, and allow us to upper bound the error of $\wh{f}$ and $\wh{\varphi}$ conditional on the design $\mathbf{X} = x$.
+In Section~\ref{sec:fixed_graph_error_bounds}, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to (functionals of) the graph $G$, and allow us to upper bound the error of $\wh{f}$ and $\wh{\varphi}$ conditional on the design $\{X_1,\ldots,X_n\} = \{x_1,\ldots,x_n\}$.
 In Sections~\ref{sec:graph_sobolev_seminorm}, \ref{sec:graph_eigenvalues}, \ref{sec:empirical_norm}, and \ref{sec:manifold} we give all our necessary probabilistic estimates on these functionals, for the particular random neighborhood graph $G = G_{n,r}$.
@@
 In Section~\ref{sec:concentration}, we state a few concentration bounds that we use repeatedly in our proofs.
+
+\paragraph{Pointwise evaluation of Sobolev functions.}
+First, however, as promised in our main text we clarify what is meant by pointwise evaluation of the regression function $f_0$.
+Strictly speaking, each $f \in H^1(X)$ is really an equivalence class, defined only up to sets of Lebesgue measure 0.
+In order to make sense of the evaluation $x \mapsto f(x)$, one must therefore pick a representative $f^{\star} \in f$.
+When $d = 1$, this is resolved in a standard way---since $H^1(\mc{X})$ embeds continuously into $C^0(\mc{X})$, there exists a continuous version of every $f \in H^1(\mc{X})$, and we take this continuous version as the representative $f^{\star}$.
+On the other hand, when $d \geq 2$, the Sobolev space $H^1(\mc{X})$ does not continuously embed into $C^0(\mc{X})$, and we must choose representatives in a different manner.
+In this case we let $f^{\star}$ be the precise representative \citep{evans15}, defined pointwise at points $x \in \mc{X}$ as
+\begin{equation*}
+f^{\star}(x) = 
+\begin{dcases*}
+\lim_{\varepsilon \to 0} \frac{1}{\nu(B(x,\varepsilon))} \int_{B(x,\varepsilon)} f(x) dx,&~~\textrm{if the limit exists,} \\
+0,&~~\textrm{otherwise.}
+\end{dcases*}
+\end{equation*}
+Note that when $d = 1$, the precise representative of any $f \in H^1(\mc{X})$ is continuous.
+
+Now we explain why the particular choice of representative is not crucial, using the notion of a Lebesgue point.
+Recall that for a locally Lebesgue integrable function $f$, a given point $x \in \mc{X}$ is a \emph{Lebesgue point} of $f$ if the limit of $1/(\nu(B(x,\varepsilon)))\int_{B(x,\varepsilon)} f(x) dx$ as $\varepsilon \to 0$ exists, and satisfies
+\begin{equation*}
+\lim_{\varepsilon \to 0} \frac{1}{\nu\bigl(B(x,\varepsilon)\bigr)} \int_{B(x,\varepsilon)} f(x) dx = f(x).
+\end{equation*}
+Let $E$ denote the set of Lebesgue points of $f$.
+By the Lebesgue differentiation theorem \citep{evans15}, if $f \in L^1(\mc{X})$ then almost every $x \in \mc{X}$ is a Lebesgue point, $\nu(\mc{X}\setminus E) = 0$.
+Since $f_0 \in H^1(\mc{X}) \subset \Leb^1(\mc{X})$, we can conclude that any function $g_0 \in f_0$ disagrees with the precise representative $f_0^{\star}$ only on a set of Lebesgue measure 0.
+Moreover, since we always assume the design distribution $P$ has a continuous density, with probability $1$ it holds that $g_0(X_i) = f_0^{\star}(X_i)$ for all $i = 1,\ldots,n$.
+This justifies the notation $f_0(X_i)$ that we use in our main text.
 
@@
 \paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}.}
-We have shown that the inequalities~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} are satisfied with probability at least $1 - \delta - C_1n\exp(-c_1nr^d)$.
-From~\eqref{eqn:neighborhood_eigenvalue} and Lemma~\ref{lem:variance_term_estimation}, we have that
-\begin{equation*}
-\sum_{k = 1}^{n}\biggl(\frac{1}{\rho \lambda_k + 1}\biggr)^{2} \geq 1 + \sum_{k = 2}^{n}\biggl(\frac{1}{C_3 \wt{\rho} k^{2/d} + 1}\biggr)^{2} \geq 1 + \frac{1}{4} \min\Bigl\{\floor{\frac{\wt{\rho}^{-d/2}}{C_3^{d/2}}} - 1 \vee 0,n\Bigr\} \geq \min\Bigl\{\frac{\wt{\rho}^{-2/d}}{8C_3^{2/d}}, \frac{n}{4} \Bigr\}
-\end{equation*}
-Therefore by Lemma~\ref{lem:ls_fixed_graph_estimation} along with~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue}, with probability at least $1 - \delta - C_1n\exp(-c_1nr^d) - \exp(-\min\{\wt{\rho}^{-d/2}/C_3^{d/2},n\}/8)$ it follows that,
+We have shown that the inequalities~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} are satisfied with probability at least $1 - \delta - C_1n\exp(-c_1nr^d)$, and throughout this proof we focus on the ``good event'' that both of these inequalities hold.
+
+Now, set $\wt{\rho} = M^{-4/(2 + d)}n^{-2/(2 + d)}$ as prescribed in Theorem~\ref{thm:laplacian_smoothing_estimation1}, and note that $\wt{\rho}^{-d/2} \leq n$ is implied by the assumption $M \leq n^{1/d}$.
+Therefore  from~\eqref{eqn:neighborhood_eigenvalue} and Lemma~\ref{lem:variance_term_estimation}, it follows that
+\begin{equation*}
+\sum_{k = 1}^{n}\biggl(\frac{1}{\rho \lambda_k + 1}\biggr)^{2} \geq 1 + \frac{1}{C_3^2}\sum_{k = 2}^{n}\biggl(\frac{1}{\wt{\rho} k^{2/d} + 1}\biggr)^{2} \geq \frac{1}{8C_3^2} \wt{\rho}^{-d/2}.
+\end{equation*}
+As a result, by Lemma~\ref{lem:ls_fixed_graph_estimation} along with~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue}, with probability at least $1 - \delta - C_1n\exp(-c_1nr^d) - \exp(-\wt{\rho}^{-d/2}/8C_3^2)$ it holds that,
 \begin{align}
 \|\wh{f} - f_0\|_n^2 & \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{n}\sum_{k = 2}^{n} \Biggl(\frac{1}{c_3 \wt{\rho}\min\{k^{2/d},r^{-2}\} + 1}\Biggr)^2 \nonumber \\
-& \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{n}\sum_{k = 1}^{n} \biggl(\frac{1}{c_3 \wt{\rho}k^{2/d} + 1}\biggr)^2 + \frac{10r^4}{c_3^2 \wt{\rho}^2}.
+& \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{nc_3^2}\sum_{k = 2}^{n} \biggl(\frac{1}{\wt{\rho}k^{2/d} + 1}\biggr)^2 + \frac{10r^4}{c_3^2 \wt{\rho}^2}.
 \label{pf:laplacian_smoothing_estimation1_1}
 \end{align}
-Setting $\wt{\rho} = M^{-4/(2 + d)}n^{-2/(2 + d)}$ as prescribed in Theorem~\ref{thm:laplacian_smoothing_estimation1},  we have from Lemma~\ref{lem:variance_term_estimation} that
-\begin{align*}
-\|\wh{f} - f_0\|_n^2 & \leq\frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{20\wt{\rho}^{-d/2}}{c_3^{d/2}n}  + \frac{10r^4}{c_3^2 \wt{\rho}^2} \\
-& = \biggl(\frac{C_2}{\delta} + \frac{20}{c_3^{d/2}}\biggr)M^{2d/(2+d)} n^{-2/(2 + d)} + \frac{10}{n} + \frac{10}{c_3^2}r^4M^{8/(2 + d)}n^{4/(2 + d)} \\
-& \leq \biggl(\frac{C_2}{\delta} + \frac{20}{c_3^{d/2}} + 10 + \frac{10}{c_3^2}\biggr)M^{2d/(2+d)} n^{-2/(2 + d)}
-\end{align*}
-with the last inequality following from~\ref{asmp:ls_kernel_radius_estimation} and the assumption $M \geq n^{-1/2}$.
-This completes the proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}.
-When $M < n^{-1/2}$, one can set $\wt{\rho} = 1$ and trace through the previous calculations to determine that $\|\wh{f} - f_0\|_n^2 \leq C/n$ for an appropriate choice of $C$.
-
-\paragraph{Nearly-optimal rates when $d = 4$.}
+The first term on the right hand side of~\eqref{pf:laplacian_smoothing_estimation1_1} is a bias term, while the second, third, and fourth terms each contribute to the variance.
+Of these, under our assumptions the third term dominates, as we show momentarily.
+First, we use Lemma~\ref{lem:variance_term_estimation} to get an upper bound on this variance term,
+\begin{equation*}
+\sum_{k = 2}^{n}\biggl(\frac{1}{\wt{\rho}k^{2/d} + 1}\biggr)^2 \leq 4\wt{\rho}^{-d/2}.
+\end{equation*}
+Then plugging this upper bound back into~\eqref{pf:laplacian_smoothing_estimation1_1}, we have that
+\begin{align*}
+\|\wh{f} - f_0\|_n^2 & \leq\frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{40\wt{\rho}^{-d/2}}{c_3^{2}n}  + \frac{10r^4}{c_3^2 \wt{\rho}^2} \\
+& = \biggl(\frac{C_2}{\delta} + \frac{40}{c_3^{2}}\biggr)M^{2d/(2+d)} n^{-2/(2 + d)} + \frac{10}{n} + \frac{10}{c_3^2}r^4M^{8/(2 + d)}n^{4/(2 + d)} \\
+& \leq \biggl(\frac{C_2}{\delta} + \frac{50}{c_3^{2}}\biggr)M^{2d/(2+d)} n^{-2/(2 + d)}
+\end{align*}
+with the last inequality following from~\ref{asmp:ls_kernel_radius_estimation} and the assumption $M \geq n^{-1/2}$. This completes the proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}.
+
+\paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_estimation2}.}
+We first establish that $\wh{f}$ achieves nearly-optimal rates when $d = 4$, and then prove the sub-optimal rates when $d > 4$.
+
+\textit{Nearly-optimal rates when $d = 4$.}
+
 Continuing on from~\eqref{pf:laplacian_smoothing_estimation1_1}, from Lemma~\ref{lem:variance_term_estimation} we have that
@@
 
-\paragraph{Suboptimal rates when $d > 4$.}
+\textit{Suboptimal rates when $d > 4$.}
+
 Once again continuing on from~\eqref{pf:laplacian_smoothing_estimation1_1}, from Lemma~\ref{lem:variance_term_estimation} we have that,
 \begin{equation*}
-\|\wh{f} - f_0\|_n^2 \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{nc_3^{d/2}\wt{\rho}^{d/2}} +  \frac{10}{n^{4/d}\wt{\rho}^2c_3^2} + \frac{10r^4}{\wt{\rho}^2c_3^2}
+\|\wh{f} - f_0\|_n^2 \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{nc_3^{2}\wt{\rho}^{d/2}} +  \frac{10}{n^{4/d}\wt{\rho}^2c_3^2} + \frac{10r^4}{\wt{\rho}^2c_3^2}
 \end{equation*}
@@
 \begin{equation*}
-\|\wh{f} - f_0\|_n^2 \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{n\wt{\rho}^{d/2}c_3^{d/2}} +  \frac{10}{n^{4/d}\wt{\rho}^2c_3^2} + \frac{10C_0^{4/d}(\log n)^{4/d}}{n^{4/d}\wt{\rho}^2c_3^2},
+\|\wh{f} - f_0\|_n^2 \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{n\wt{\rho}^{d/2}c_3^{2}} +  \frac{10}{n^{4/d}\wt{\rho}^2c_3^2} + \frac{10C_0^{4/d}(\log n)^{4/d}}{n^{4/d}\wt{\rho}^2c_3^2},
 \end{equation*}
@@
 \begin{equation*}
-\|\wh{f} - f_0\|_n^2 \leq \biggl(\frac{C_2}{\delta} + \frac{10}{c_3^2} + \frac{10C_0^{4/d}}{c_3^2}\biggr)M^{4/d}\biggl(\frac{\log(n)}{n^{1/3}}\biggr)^{4/d} + \frac{10}{c_3^{d/2}}M^{d/3}n^{-1/3} + \frac{10}{n}.
+\|\wh{f} - f_0\|_n^2 \leq \biggl(\frac{C_2}{\delta} + \frac{10}{c_3^2} + \frac{10C_0^{4/d}}{c_3^2}\biggr)M^{4/3}\biggl(\frac{\log n}{n^{1/3}}\biggr)^{4/d} + \frac{10}{c_3^{d/2}}M^{d/3}n^{-1/3} + \frac{10}{n}.
 \end{equation*}
@@
 	\label{lem:variance_term_estimation}
-	For any $t > 0$,
-	\begin{equation*}
-	\frac{1}{4}\min\Bigl\{\floor{t^{-d/2}} - 1 \vee 0,n\Bigr\} \leq \sum_{k = 2}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^2 \leq t^{-d/2} +
+	For any $t > 0$ such that $1 \leq t^{-d/2} \leq n$,
+	\begin{equation*}
+	\frac{1}{8}t^{-d/2} - 1 \leq \sum_{k = 2}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^2 \leq t^{-d/2} +
 	\begin{cases*}
-	t^{-d/2},& ~~\textrm{if $d < 4$} \\
+	3t^{-d/2},& ~~\textrm{if $d < 4$} \\
 	\frac{1}{t^2}\log n,& ~~\textrm{if $d = 4$} \\
-	\frac{1}{t^2}n^{-(4/d - 1)},&~~\textrm{if $d > 4$.}
+	\frac{1}{t^2}n^{1 - 4/d},&~~\textrm{if $d > 4$.}
 	\end{cases*}
@@
 \paragraph{Proof of Lemma~\ref{lem:variance_term_estimation}.}
-We begin by proving the upper bounds. Observe that
-\begin{align*}
- \sum_{k = 1}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^2 & \leq t^{-d/2} + \frac{1}{t^2}\sum_{k = t^{-d/2} + 1}^{n} \frac{1}{k^{4/d}} \\
- & \leq t^{-d/2} + \frac{1}{t^2} \int_{t^{-d/2}}^{n} x^{-4/d} \,dx
-\end{align*}
-and the various upper bounds (for $d < 4$, $d = 4$, and $d > 4$) follow upon computing the integral.
-
-For the lower bound, we simply recognize that for each $k = 2,\ldots,n$ such that $k \leq \floor{t^{-d/2}}$, it holds that $1/(tk^{2/d} + 1)^2 \geq 1/4$, and there are at least $\min\Bigl\{\floor{t^{-d/2}} - 1 \vee 0,n\Bigr\}$ such values of $k$.
+We begin by proving the upper bounds. Treating the sum over $k$ as a Riemann sum of a non-increasing function, we have that
+\begin{align*}
+ \sum_{k = 2}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^2 & \leq \int_{1}^{n} \biggl(\frac{1}{tx^{2/d} + 1}\biggr)^2 \,dx \leq t^{-d/2} + \int_{t^{-d/2}}^{n} \biggl(\frac{1}{tx^{2/d} + 1}\biggr)^2 \,dx \leq t^{-d/2} + \frac{1}{t^2} \int_{t^{-d/2}}^{n} x^{-4/d} \,dx.
+\end{align*}
+The various upper bounds (for $d < 4$, $d = 4$, and $d > 4$) then follow upon computing the integral.
+
+For the lower bound, we simply recognize that for each $k = 2,\ldots,n$ such that $k \leq \floor{t^{-d/2}}$, it holds that $1/(tk^{2/d} + 1)^2 \geq 1/4$, and there are at least $\min\Bigl\{\floor{t^{-d/2}} - 1,n - 1\Bigr\} > \frac{1}{2}t^{-d/2} - 1$ such values of $k$.
 

### 2021-02-24 manuscripts/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
AJG: Updated draft of Laplacian smoothing AISTATS paper, in response to Ryan's comments.

@@
 
-	\aistatsaddress{Carnegie Mellon University \And Carnegie Mellon University \And Carnegie Mellon University}]
+  \vspace{4pt}
+
+	\aistatsaddress{Carnegie Mellon University}]
 
@@
 We leave analysis of the Sobolev case to future work.
-%	\item Under the stronger assumption that $f_0$ is Lipschitz smooth, we can replace the factor of $\delta$ in the high probability bound by the sharper $\delta^2/n$.
+	\item Additionally, under the stronger assumption that $f_0$ is Lipschitz smooth, we can replace the factor of $\delta$ in the high probability bound by the sharper $\delta^2/n$.
 \end{itemize}
@@
 \end{enumerate}
-
-Now define
+We will also require that the radius of the graph Sobolev class not be too large. Precisely, we will require $M \leq M_{\max}(d)$, where we define
 \begin{equation*}
-M_{\max}(d) =
+M_{\max}(d) :=
 \begin{cases*}
@@
 \end{equation*}
-
-Note that when $d < 4$ and $M \leq M_{\max}(d)$, there always exists a choice of $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}. 
 We now state Theorem~\ref{thm:laplacian_smoothing_testing}, our main testing result.
@@
   Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ with $d < 4$, and where $M \leq M_{\max}(d)$.
-Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume \smash{$G_{n,r}$} is computed with a kernel $K$ satifsying \ref{asmp:kernel}.
+Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume \smash{$G_{n,r}$} is computed with a kernel $K$ satisfying \ref{asmp:kernel}.
 There are constants $N,C,c>0$ such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_testing}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$}, satisfies the following: for all $b \geq 1$, and all $f_0$ such that
@@
 	\label{thm:laplacian_smoothing_estimation_manifold}
-  As in Theorem \ref{thm:laplacian_smoothing_estimation1}, but where $\Xset \subseteq \Rd$ is now a manifold with intrinsic dimension $m < 4$, and the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}.
-There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_estimation_manifold}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + m)} (nr^{m + 2})^{-1} n^{-2/(2 + m)}$}, satisfies
+  As in Theorem \ref{thm:laplacian_smoothing_estimation1}, but where $\Xset \subseteq \Rd$ is now a manifold with intrinsic dimension $m < 4$, the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}, and $M \leq n^{1/m}$.
+There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any \smash{$n \geq N$}, and any radius $r$ as in \ref{asmp:ls_kernel_radius_estimation_manifold}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + m)} (nr^{m + 2})^{-1} n^{-2/(2 + m)}$}, satisfies
 	\begin{equation*}
@@
 	\end{equation*}
-  with probability at least $1 - \delta -  C_1 n\exp(-cnr^m) - \exp(-c \min\{M^{m/(2m + 4)} n^{m/(2+m)},n\})$.
+  with probability at least $1 - \delta -  C_1 n\exp(-cnr^m) - \exp(-c(M^2n)^{m/(2+m)})$.
 % RJT: Alden, please verify that what I wrote is correct.
+% AJG: Verified.
 \end{theorem}
@@
 % RJT: Alden, please verify that what I wrote is correct.
+% AJG: Verified.
 \end{theorem}

### 2021-02-24 manuscripts/laplacian_smoothing/AISTATS/supplement_minimax_optimal_laplacian_smoothing.tex
AJG: Updated draft of Laplacian smoothing AISTATS paper, in response to Ryan's comments.

@@
 
+\runningtitle{}
+
 \onecolumn	
@@
 
-To prove Lemma~\ref{lem:graph_sobolev_seminorm}, we will show that for any $f \in H^1(\Xset)$,
-\begin{equation*}
-\Ebb\Bigl[f^{\top} \Lap_{n,r} f\Bigr] \leq C_2 n^2 r^{d + 2} \|f\|_{H^1(\Xset)}^2,
-\end{equation*}
-whence the claim follows immediately by Markov's inequality (recall that $\Lap_{n,r}$ is positive semi-definite, and therefore $f^{\top} \Lap_{n,r} f$ is a non-negative random variable).
+Throughout this proof, we will assume that $f \in H^1(\Xset)$ has zero-mean, meaning $\int_{\Xset} f(x) \,dx = 0$.
+This is without loss of generality---assuming for the moment that~\eqref{eqn:graph_sobolev_seminorm} holds for zero-mean functions, for any $f \in H^1(\Xset)$, taking $a = \int_{\Xset} f(x) \,dx$ and $g = f - a$, we have that
+\begin{equation*}
+f^{\top} L_{n,r} f = g^{\top} L_{n,r} g \leq \frac{C_2}{\delta} n^2 r^{d + 2} |g|_{H^1(\mc{X})}^2 = \frac{C_2}{\delta} n^2 r^{d + 2} |f|_{H^1(\mc{X})}^2.
+\end{equation*} 
+Now, for any zero-mean function $f \in H^1(\mc{X})$ it follows by the Poincare inequality (see Section 5.8, Theorem 1 of \citet{evans10}) that $\|f\|_{H^1(\mc{X})}^2 \leq C_8 |f|_{H^1(\mc{X})}^2$, for some constant $C_8$ that does not depend on $f$.
+Therefore, to prove Lemma~\ref{lem:graph_sobolev_seminorm}, it suffices to show that
+\begin{equation*}
+\Ebb\Bigl[f^{\top} \Lap_{n,r} f\Bigr] \leq C n^2 r^{d + 2} \|f\|_{H^1(\Xset)}^2,
+\end{equation*}
+since the high-probability upper bound then follows immediately by Markov's inequality.
+(Recall that $\Lap_{n,r}$ is positive semi-definite, and therefore $f^{\top} \Lap_{n,r} f$ is a non-negative random variable).
 
@@
 
-Take $\Omega$ to be an arbitrary bounded open set such that $B(x,r) \subset \Omega$ for all $x \in \mc{X}$.
+Now, take $\Omega$ to be an arbitrary bounded open set such that $B(x,c_0) \subset \Omega$ for all $x \in \mc{X}$.
 For the remainder of this proof, we will assume that (i) $f \in H^1(\Omega)$ and additionally (ii) $\|f\|_{H^1(\Omega)} \leq C_5 \|f\|_{H^1(\mc{X})}$ for a constant $C_5$ that does not depend on $f$.
-This is without loss of generality, since by Theorem~1 of~\citet{evans10} there exists an extension operator $E: H^1(\mc{X}) \to H^1(\Omega)$ for which the extension $Ef$ satisfies both (i) and (ii).
+This is without loss of generality, since by Theorem~1 in Chapter 5.4 of~\citet{evans10} there exists an extension operator $E: H^1(\mc{X}) \to H^1(\Omega)$ for which the extension $Ef$ satisfies both (i) and (ii).
 Additionally, we will assume $f \in C^{\infty}(\Omega)$.
@@
 \end{equation*}
-proving the claim of Lemma~\ref{lem:graph_sobolev_seminorm} upon taking $C_2 := C_5 \sigma_K p_{\max}^2$ in the statement of the lemma.
+proving the claim of Lemma~\ref{lem:graph_sobolev_seminorm} upon taking $C_2 := C_8C_5 \sigma_K p_{\max}^2$ in the statement of the lemma.
 
@@
 \paragraph{Proof of Lemma~\ref{lem:empirical_norm_sobolev}.}
-To establish~\eqref{eqn:empirical_norm_sobolev}, it is sufficient to show that
+In this proof, we will find it more convenient to deal with the parameterization $b = 1/\delta$. To establish~\eqref{eqn:empirical_norm_sobolev}, it is sufficient to show that
 \begin{equation*}
@@
 \begin{equation*}
-\mathbb{E}\bigl[\norm{f}_n^4\bigr] = \frac{(n-1)}{n}\left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + \frac{\mathbb{E}\Bigl[\bigl(f(x_1)\bigr)^4\Bigr]}{n} \leq \left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + p_{\max}\frac{\norm{f}_{\Leb^4(\Xset)}^4}{n}.
+\mathbb{E}\bigl[\norm{f}_n^4\bigr] = \frac{(n-1)}{n}\left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + \frac{\mathbb{E}\Bigl[\bigl(f(X_1)\bigr)^4\Bigr]}{n} \leq \left(\mathbb{E}\Bigl[\norm{f}_n^2\Bigr]\right)^2 + p_{\max}\frac{\norm{f}_{\Leb^4(\Xset)}^4}{n}.
 \end{equation*}
@@
 \paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}.}
-We have shown that the inequalities~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} are satisfied with probability at least $1 - \delta - C_1n\exp(-c_1nr^d)$, and throughout this proof we focus on the ``good event'' that both of these inequalities hold.
+We have shown that the inequalities~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} are satisfied with probability at least $1 - \delta - C_1n\exp(-c_1nr^d)$, and throughout this proof we take as granted that both of these inequalities hold.
 
@@
 \paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_testing}.}
-Recall that we have shown that the inequalities~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} are satisfied with probability at least $1 - \delta - C_1n\exp(-c_1nr^d)$.
+Let $\delta = 1/b$.
+Recall that we have shown that the inequalities~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} are satisfied with probability at least $1 - 1/b - C_1n\exp(-c_1nr^d)$, and throughout this proof we take as granted that both of these inequalities hold.
 
@@
 First, we upper bound the right hand side of this inequality.
-By~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue},
-\begin{align*}
-\frac{2 \rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{4b}{n} \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{1/2} & \leq C_2b\wt{\rho}M^2 + \frac{4b}{n}\Biggl[1 + \biggl(\sum_{k = 2}^{n} \frac{1}{(c_3\wt{\rho}k^{2/d} + 1)^4} \biggr)^{1/2} + \frac{r^4n^{1/2}}{c_3^2\wt{\rho}^2}\Biggr] \\
-& \leq C_2b\wt{\rho}M^2 + \frac{4b}{n}\biggl(1 + \frac{\sqrt{2}}{c_3^{d/4}}\wt{\rho}^{-d/4} + \frac{r^4n^{1/2}}{c_3^2\wt{\rho}^2}\biggr) \\ 
-& \leq \Bigl(C_2 + 4 + \frac{4\sqrt{2}}{c_3^{d/4}} + \frac{4}{c_3^2}\Bigr)bM^{2d/(4 + d)} n^{-4/(4 + d)}
-\end{align*}
-where the second inequality is justfied by Lemma~\ref{lem:variance_term_testing}, and in the third inequality we set $\wt{\rho} = M^{-8/(4 + d)}n^{-4/(4 + d)}$ as prescribed by Theorem~\ref{thm:laplacian_smoothing_testing}, and use both the upper bound on $r$ assumed in~\ref{asmp:ls_kernel_radius_testing} as well as the fact that $M \geq n^{-1/2}$.
+Set $\wt{\rho} = M^{-8/(4 + d)}n^{-4/(4 + d)}$ as prescribed by Theorem~\ref{thm:laplacian_smoothing_testing}, it follows from~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} that
+\begin{align*}
+\frac{2 \rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{4b}{n} \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{1/2} & \leq C_2b\wt{\rho}M^2 + \frac{4b}{n}\Biggl[1 + \frac{1}{c_3^2}\biggl(\sum_{k = 2}^{n} \frac{1}{(\wt{\rho}k^{2/d} + 1)^4} \biggr)^{1/2} + \frac{r^4n^{1/2}}{c_3^2\wt{\rho}^2}\Biggr] \\
+& \leq C_2b\wt{\rho}M^2 + \frac{4b}{n}\biggl(1 + \frac{\sqrt{2}}{c_3^{2}}\wt{\rho}^{-d/4} + \frac{r^4n^{1/2}}{c_3^2\wt{\rho}^2}\biggr) \\ 
+& \leq \Bigl(C_2 + 4 + \frac{4\sqrt{2}}{c_3^{2}} + \frac{4}{c_3^2}\Bigr)bM^{2d/(4 + d)} n^{-4/(4 + d)}.
+\end{align*}
+The second inequality in the above is justified by Lemma~\ref{lem:variance_term_testing}, keeping in mind that $M \leq M_{\max}(d)$ implies that $\wt{\rho}^{-d/2} \leq n$.
+The third inequality follows from the upper bound on $r$ assumed in~\ref{asmp:ls_kernel_radius_testing} as well as the fact that $M \geq n^{-1/2}$.
 
@@
 Recall that by assumption, $M \leq M_{\max}(d)$.
-Therefore, taking $C \geq C_6$ in~\eqref{eqn:laplacian_smoothing_testing}, we have that~\eqref{eqn:empirical_norm_sobolev_1} is satisfied, so that it follows from~\eqref{eqn:empirical_norm_sobolev} that
-\begin{equation*}
-\|f\|_n^2 \geq \frac{\|f\|_{P}^2}{b} \geq \frac{p_{\min}}{b} \|f\|_{\Leb^2(\Xset)}^2 \geq CbM^{2d/(4 + d)}n^{-4/(4 + d)},
+Therefore, taking $C \geq C_6$ in~\eqref{eqn:laplacian_smoothing_testing}, implies that the lower bound on $\|f\|_{\Leb^2(\mc{X})}$ in~\eqref{eqn:empirical_norm_sobolev_1} is satisfied.
+As a result, it follows from~\eqref{eqn:empirical_norm_sobolev} that
+\begin{equation*}
+\|f\|_n^2 \geq \frac{\mathbb{E}[\|f\|_n^2]}{b} \geq \frac{p_{\min}}{b} \|f\|_{\Leb^2(\Xset)}^2 \geq CbM^{2d/(4 + d)}n^{-4/(4 + d)},
 \end{equation*}
 with probability at least $1 - 5/b$.
-Taking $C \geq C_2 + 4 + (4\sqrt{2})/c_3^{d/4} + 4/c_3^2$ in~\eqref{eqn:laplacian_smoothing_testing} thus implies~\eqref{eqn:ls_fixed_graph_testing_critical_radius}, and we may therefore use Lemma~\ref{lem:ls_fixed_graph_testing} to upper bound the type II error the Laplacian smoothing test $\wh{\varphi}$.
+Taking $C \geq C_2 + 4 + (4\sqrt{2})/c_3^{2} + 4/c_3^2$ in~\eqref{eqn:laplacian_smoothing_testing} thus implies~\eqref{eqn:ls_fixed_graph_testing_critical_radius}, and we may therefore use Lemma~\ref{lem:ls_fixed_graph_testing} to upper bound the type II error the Laplacian smoothing test $\wh{\varphi}$.
 Observe that by~\eqref{eqn:neighborhood_eigenvalue} and the lower bound in Lemma~\ref{lem:variance_term_testing},
 \begin{equation*}
-\sum_{k = 1}^{n} \biggl(\frac{1}{\rho \lambda_k + 1}\biggr)^4 \geq 1 + \sum_{k = 2}^{n}\biggl(\frac{1}{C_3\wt{\rho}k^{2/d} + 1}\biggr)^4 \geq 1 + \frac{1}{16}\min\biggl\{\floor{\frac{\wt{\rho}^{-d/2}}{C_3^{d/2}}} - 1 \vee 0, n \biggr\} \geq \min\biggl\{\frac{\wt{\rho}^{-d/2}}{32 C_3^{d/2}}, \frac{n}{16} \biggr\}.
+\sum_{k = 1}^{n} \biggl(\frac{1}{\rho \lambda_k + 1}\biggr)^4 \geq 1 + \frac{1}{C_3^4}\sum_{k = 2}^{n}\biggl(\frac{1}{\wt{\rho}k^{2/d} + 1}\biggr)^4 \geq \frac{1}{32C_3^4}\wt{\rho}^{-d/2}.
 \end{equation*}
@@
 \Pbb_{f_0}\bigl(\wh{T} \leq \wh{t}_b\bigr) & \leq \frac{6}{b} + \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2} + C_1n\exp(-c_1nr^d)\\
-& \leq \frac{10}{b} + \frac{32\sqrt{2}}{b} \max\Bigl\{ C_3^{d/4}\wt{\rho}^{d/4} ,n^{-1/2}\Bigr\} + C_1n\exp(-c_1nr^d),
+& \leq \frac{10}{b} + \frac{32\sqrt{2}}{b} C_3^{2}\wt{\rho}^{d/4} + C_1n\exp(-c_1nr^d),
 \end{align*}
@@
 	\label{lem:variance_term_testing}
-	For any $t > 0$, when $d < 4$,
-	\begin{equation*}
-	\frac{1}{16}\min\Bigl\{\floor{t^{-d/2}} - 1 \vee 0,n\Bigr\} \leq \sum_{k = 2}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^4 \leq 2t^{-d/2}.
+	Suppose $d \leq 4$. Then for any $t > 0$ such that $1 \leq t^{-d/2} \leq n$,
+	\begin{equation*}
+	\frac{1}{32}t^{-d/2} - 1 \leq \sum_{k = 2}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^4 \leq 2t^{-d/2}.
 	\end{equation*}
@@
 \begin{align*}
-\sum_{k = 1}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^4 & \leq t^{-d/2} + \frac{1}{t^4}\sum_{k = t^{-d/2} + 1}^{n} \frac{1}{k^{8/d}} \\
-& \leq t^{-d/2} + \frac{1}{t^4} \int_{t^{-d/2}}^{n} x^{-8/d} \,dx \\
-& \leq 2t^{-d/2}.
-\end{align*}
-
+\sum_{k = 1}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^4 \leq t^{-d/2} + \frac{1}{t^4}\sum_{k = t^{-d/2} + 1}^{n} \frac{1}{k^{8/d}} \leq t^{-d/2} + \frac{1}{t^4} \int_{t^{-d/2}}^{n} x^{-8/d} \,dx \leq 2t^{-d/2}.
+\end{align*}
 The lower bound follows from the same logic as we used to derive the lower bound in Lemma~\ref{lem:variance_term_estimation}.

### 2021-02-24 code/plot.R
AJG: Changed figures to avoid cutoff on the left hand side.

@@ -3,7 +3,7 @@ library(ggplot2)
 library(gridExtra)
 
 ### Plot. ###
-save_directory <- "data/AISTATS/cosine/d1_largesample" # Please change this to whichever directory you prefer.
+save_directory <- "data/AISTATS/cosine/d5_largesample" # Please change this to whichever directory you prefer.
 plot_directory <- file.path(save_directory,"plots")
 dir.create(plot_directory)
 
@@ -83,6 +83,7 @@ if(d == 1)
 
 # Plot of mse---for best choice of tuning parameter---by sample size
 pdf(file.path(plot_directory,"mse_by_sample_size.pdf"))
+par(mar = c(5.1,4.5,4.1,4.1)) # The left hand side was getting cut off a tad.
 plot_dfs_best_mse <- vector(mode = "list", length = length(methods))
 names(plot_dfs_best_mse) <- names(methods)
 fitted_slopes <- numeric(length(methods))

### 2021-02-25 manuscripts/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
Added one line
Corrected a few minor typos for AISTATS camera-ready version.

@@
 These upper bounds match the minimax optimal estimation and testing rates of convergence over the first-order Sobolev class $H^1(\Xset)$, for $\Xset \subseteq \Rd$ and $1 \leq d < 4$; in the estimation problem, for $d = 4$, they are optimal modulo a $\log n$ factor.
-Additionally, we prove that Laplacian smoothing is manifold-adaptive: if $\Xset \subseteq \R^d$ is an $m$-dimensional manifold with $m < d$, then the error rate of Laplacian smoothing (in either estimation or testing) depends only on $m$, in the same way it would if $\Xset$ were a full-dimensional set in $\Reals^m$.
+Additionally, we prove that Laplacian smoothing is manifold-adaptive: if $\Xset \subseteq \Reals^d$ is an $m$-dimensional manifold with $m < d$, then the error rate of Laplacian smoothing (in either estimation or testing) depends only on $m$, in the same way it would if $\Xset$ were a full-dimensional set in $\Reals^m$.
 \end{abstract}
@@
 	\item \emph{Fixed design perspective.} Here one treats the design points $X_1,\ldots,X_n$ and the graph $G$ as fixed, and carries out inference on $f_0(X_i)$, $i=1,\ldots,n$.
-In this problem setting, tight upper bounds have been derived on the error of various graph-based methods (e.g., \citet{wang2016,hutter2016,sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}) and tests (e.g., \citet{sharpnack2010identifying,sharpnack2013b,sharpnack2013,sharpnack2015}), which certify that such procedures are \emph{optimal} over ``function'' classes (in parentheses because these classes really model the $n$-dimensional vector of evaluations).
+In this problem setting, tight upper bounds have been derived on the error of various graph-based methods (e.g., \citet{wang2016,hutter2016,sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}) and tests (e.g., \citet{sharpnack2010identifying,sharpnack2013b,sharpnack2013,sharpnack2015}), which certify that such procedures are \emph{optimal} over ``function'' classes (in quotes because these classes really model the $n$-dimensional vector of evaluations).
 The upside of this work is its generality: in this setting $G$ need not be a geometric graph, but in principle it could be any graph over $V=\{1,\ldots,n\}$.
-The downside is that, in the context of nonparametric regression, it is as not natural to think of the evaluations of $f_0$ as exhibiting smoothness over some fixed pre-defined graph $G$, and more natural to speak of the smoothness of the function $f_0$ itself.
+The downside is that, in the context of nonparametric regression, it is not as natural to think of the evaluations of $f_0$ as exhibiting smoothness over some fixed pre-defined graph $G$, and more natural to speak of the smoothness of the function $f_0$ itself.
 	\item \emph{Random design perspective.} Here one treats the design points $X_1,\ldots,X_n$ as independent samples from some distribution $P$ supported on a domain $\Xset \subseteq \Rd$.
@@
 In this paper, we will adopt the random design perspective, and seek to answer the following question:
-\begin{quote}{When $f_0$ is assumed to smooth in a continuum sense, does Laplacian smoothing achieve optimal performance for estimation and goodness-of-fit testing?} 
+\begin{quote}{When $f_0$ is assumed to be smooth in a continuum sense, does Laplacian smoothing achieve optimal performance for estimation and goodness-of-fit testing?} 
 \end{quote}
@@
 \end{equation*}
-i.e., $f$ achieves perfect (zero) data loss and arbitrarily small penalty in the problem \eqref{eqn:thin_plate_spline}.
+In other words, $f$ achieves perfect (zero) data loss and arbitrarily small penalty in the problem \eqref{eqn:thin_plate_spline}.
 This will clearly not lead to a consistent estimator of $f_0$ across the design points (as it always yields $Y_i$ at each $X_i$).
@@
 It's fine, I wouldn't change anything because it would involve explaining a lot more, I'm just pointing this out.
-(Throughout we assume $M \geq n^{-1/2}$---as otherwise the minimax rate is just the parametric rate $n^{-1}$.)
+(Throughout we assume $M \geq n^{-1/2}$, as otherwise the minimax rate is just the parametric rate $n^{-1}$.)
 
@@
 We leave analysis of the Sobolev case to future work.
-	\item Additionally, under the stronger assumption that $f_0$ is Lipschitz smooth, we can replace the factor of $\delta$ in the high probability bound by the sharper $\delta^2/n$.
+	\item Under the assumption that $f_0$ is Lipschitz smooth, we can also replace the factor of $\delta$ in the high probability bound by the sharper $\delta^2/n$.
 \end{itemize}
@@
 	\end{equation*}
+again with the same probability guarantee.
 % RJT: Alden, please verify what I wrote is correct, and fill in the result for d \geq 5.
@@
 
-Of course, we know this is not the case: $\wh{f}$ and $\wt{f}$ look very differently when $d > 1$.
+Of course, we know this is not the case: $\wh{f}$ and $\wt{f}$ look very different when $d > 1$.
 What is driving this difference?
@@
 Even among continuous and differentiable functions $f$, as we explained previously, we can use ``bump'' functions (as in \citet{green93}) to construct $f$ that interpolates the pairs $(X_i,Y_i)$, $i=1,\ldots,n$ and achieves arbitrarily small penalty (and hence criterion) in \eqref{eqn:thin_plate_spline}.
-In this sense, any estimator resulting from solving \eqref{eqn:thin_plate_spline} estimator will clearly be inconsistent.
+In this sense, any estimator resulting from solving \eqref{eqn:thin_plate_spline} will clearly be inconsistent.
 

### 2021-02-28 manuscripts/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
Typos and copyedits

--- 2021-02-24 manuscripts/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
@@
 	\label{thm:laplacian_smoothing_estimation1}
-  Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ has dimension $d < 4$ and where $M \leq n^{1/d}$.
+  Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ has dimension $d < 4$ and $M \leq n^{1/d}$.
 Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume the neighborhood graph \smash{$G_{n,r}$} is computed with a kernel $K$ satifsying \ref{asmp:kernel}.
 There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_estimation}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$}, satisfies
-	\begin{equation}
-	\label{eqn:laplacian_smoothing_estimation1}
+	\begin{equation*}
 	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(2 + d)} n^{-2/(2 + d)},
-	\end{equation}
+	\end{equation*}
   with probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c (M^2n)^{d/(2 + d)})$.
@@
 % AJG: Added.
-	\item The assumption $M \leq n^{1/d}$ guarantees that the bound in~\eqref{eqn:laplacian_smoothing_estimation1} is meaningful, i.e.
-that the right hand side of~\eqref{eqn:laplacian_smoothing_estimation1} is of at most a constant order.
+	\item The assumption $M \leq n^{1/d}$ ensures that the upper bound provided in the theorem is meaningful (i.e., ensures it is of most a constant order).
 	\item The lower bound on $r$ imposed in condition \ref{asmp:ls_kernel_radius_estimation} is compatible with practice, where by far the most common choice of radius is the connectivity threshold \smash{$r \asymp (\log(n)/n)^{1/d}$}, which makes \smash{$G_{n,r}$} as sparse as possible while still being connected, for maximum computational efficiency.
@@
 	\label{thm:laplacian_smoothing_estimation2}
-  Under the assumptions of Theorem~\ref{thm:laplacian_smoothing_estimation1}, if $\Xset$ has dimension $d = 4$, \smash{$r \asymp (\log n/n)^{1/4}$} and \smash{$\rho = M^{-2/3}(nr^6)^{-1}(\log n/n)^{1/3}$}, then we obtain
+  Under the assumptions of Theorem~\ref{thm:laplacian_smoothing_estimation1}, if instead $\Xset$ has dimension $d = 4$, \smash{$r \asymp (\log n/n)^{1/4}$} and \smash{$\rho = M^{-2/3}(nr^6)^{-1}(\log n/n)^{1/3}$}, then we obtain
 	\begin{equation*}
@@
 	\label{thm:laplacian_smoothing_estimation_manifold}
-  As in Theorem \ref{thm:laplacian_smoothing_estimation1}, but where $\Xset \subseteq \Rd$ is now a manifold with intrinsic dimension $m < 4$, the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}, and $M \leq n^{1/m}$.
-There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any \smash{$n \geq N$}, and any radius $r$ as in \ref{asmp:ls_kernel_radius_estimation_manifold}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + m)} (nr^{m + 2})^{-1} n^{-2/(2 + m)}$}, satisfies
+  As in Theorem \ref{thm:laplacian_smoothing_estimation1}, but where $\Xset \subseteq \Rd$ is a manifold with intrinsic dimension $m < 4$, the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}, and $M \leq n^{1/m}$.
+There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any \smash{$n \geq N$}, and any $r$ as in \ref{asmp:ls_kernel_radius_estimation_manifold}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + m)} (nr^{m + 2})^{-1} n^{-2/(2 + m)}$}, satisfies
 	\begin{equation*}
@@
 We conclude by mentioning a couple more.
-In practice, it is more common to use $k$-nearest-neighbor (kNN) graphs than neighborhood graphs, due to the guaranteed connectivity and sparsity of the former; we suspect that by building on the work of \citet{calder2019}, one can show that our main results all hold under the kNN graph as well.
-In another direction, one can also generalize Laplacian smoothing by replacing the penalty \smash{$f^\top \Lap_{n,r} f$} with \smash{$f^\top \Lap_{n,r}^k f$} for $k > 1$.
+In practice, it is more common to use a $k$-nearest-neighbor (kNN) graph than a neighborhood graph, due to the guaranteed connectivity and sparsity of the former; we suspect that by building on the work of \citet{calder2019}, one can show that our main results all hold under the kNN graph as well.
+In another direction, one can also generalize Laplacian smoothing by replacing the penalty \smash{$f^\top \Lap_{n,r} f$} with \smash{$f^\top \Lap_{n,r}^k f$}, for an integer $k > 1$.
 The hope is that this would then achieve minimax optimal rates over the higher-order Sobolev class $H^k(\Xset)$.

### 2021-02-28 manuscripts/laplacian_smoothing/AISTATS/supplement_minimax_optimal_laplacian_smoothing.tex
Typos and copyedits

@@
 \onecolumn	
-\aistatstitle{Supplementary Materials for ``Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs''}
+\aistatstitle{Supplementary Material for ``Minimax Optimal Regression over
+  Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs''}
+
+\aistatsauthor{Alden Green \And Sivaraman Balakrishnan \And Ryan J. Tibshirani}
+
+\vspace{4pt}
+
+\aistatsaddress{Carnegie Mellon University}
 
@@
 By the Lebesgue differentiation theorem \citep{evans15}, if $f \in L^1(\mc{X})$ then almost every $x \in \mc{X}$ is a Lebesgue point, $\nu(\mc{X}\setminus E) = 0$.
-Since $f_0 \in H^1(\mc{X}) \subset \Leb^1(\mc{X})$, we can conclude that any function $g_0 \in f_0$ disagrees with the precise representative $f_0^{\star}$ only on a set of Lebesgue measure 0.
+Since $f_0 \in H^1(\mc{X}) \subseteq \Leb^1(\mc{X})$, we can conclude that any function $g_0 \in f_0$ disagrees with the precise representative $f_0^{\star}$ only on a set of Lebesgue measure 0.
 Moreover, since we always assume the design distribution $P$ has a continuous density, with probability $1$ it holds that $g_0(X_i) = f_0^{\star}(X_i)$ for all $i = 1,\ldots,n$.
-This justifies the notation $f_0(X_i)$ that we use in our main text.
+This justifies the notation $f_0(X_i)$ used in the main text.
 
@@
 \label{sec:fixed_graph_error_bounds}
+
 In this section, we adopt the fixed design perspective; or equivalently, condition on $X_i = x_i$ for $i = 1,\ldots,n$.
@@
 
-Recall~\eqref{eqn:laplacian_smoothing} and~\eqref{eqn:laplacian_smoothing_test}: the Laplacian smoothing estimator of $f_0$ on $G$ is
+Recall \eqref{eqn:laplacian_smoothing} and \eqref{eqn:laplacian_smoothing_test}: the Laplacian smoothing estimator of $f_0$ on $G$ is
 \begin{equation*}
@@
 \subsection{Error bounds for linear smoothers}
+
 Let $S \in \Reals^{n \times n}$ be a square, symmetric matrix, and let 
@@
 
-\textit{Variance.} Starting from~\eqref{pf:linear_smoother_fixed_graph_testing1} and recalling $\Var(Z_k^2) = 2$, we derive
+\textit{Variance.} Starting from \eqref{pf:linear_smoother_fixed_graph_testing1} and recalling $\Var(Z_k^2) = 2$, we derive
 \begin{equation}
@@
 
-\textit{Bounding Type I and Type II error.} The upper bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeI} follows immediately from~\eqref{pf:linear_smoother_fixed_graph_testing_mean}, ~\eqref{pf:linear_smoother_fixed_graph_testing_var}, and Chebyshev's inequality.
-
-The bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeII} also follows from Chebyshev's inequality, as can be seen by the following manipulations,
+\textit{Bounding Type I and Type II error.} The upper bound \eqref{eqn:linear_smoother_fixed_graph_testing_typeI} follows immediately from \eqref{pf:linear_smoother_fixed_graph_testing_mean}, \eqref{pf:linear_smoother_fixed_graph_testing_var}, and Chebyshev's inequality.
+
+The bound \eqref{eqn:linear_smoother_fixed_graph_testing_typeII} also follows from Chebyshev's inequality, as can be seen by the following manipulations,
 \begin{equation*}
@@
 \end{equation*}
-In the previous expression, $(i)$ and $(ii)$ follow since assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and equation~\eqref{pf:linear_smoother_fixed_graph_testing_mean} together imply
+In the previous expression, $(i)$ and $(ii)$ follow since assumption \eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and equation \eqref{pf:linear_smoother_fixed_graph_testing_mean} together imply
 \begin{equation*}
@@
 \end{equation*}
-$(iii)$ follows from assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and the inequality~\eqref{pf:linear_smoother_fixed_graph_testing_var}, and $(iv)$ follows from~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}.
-
-\subsection{Analysis of Laplacian Smoothing}
+$(iii)$ follows from assumption \eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and the inequality \eqref{pf:linear_smoother_fixed_graph_testing_var}, and $(iv)$ follows from \eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}.
+
+\subsection{Analysis of Laplacian smoothing}
+
 Upper bounds on the mean squared error of $\wh{f}$, and Type I and Type II error of $\wh{T}$, follow from setting $S = (\rho L + I)^{-1}$ in Lemmas~\ref{lem:linear_smoother_fixed_graph_estimation} and~\ref{lem:linear_smoother_fixed_graph_testing}.
@@
 \end{lemma}
+
 \paragraph{Proof of Lemma~\ref{lem:ls_fixed_graph_estimation}.}
@@
 Let $\wh{S} := (\rho \Lap + \Id)^{-1}$, and note that $\wh{T} = \frac{1}{n}{\bf Y}^{\top} \wh{S}^2 {\bf Y}$.
-The bound on Type I error~\eqref{eqn:ls_fixed_graph_testing_typeI} follows immediately from \eqref{eqn:linear_smoother_fixed_graph_testing_typeI}.
-To establish the bound on Type II error, we must lower bound $f_0^{\top} \wh{S}^2 f_0$. We first note that by assumption~\eqref{eqn:ls_fixed_graph_testing_critical_radius},
+The bound on Type I error \eqref{eqn:ls_fixed_graph_testing_typeI} follows immediately from \eqref{eqn:linear_smoother_fixed_graph_testing_typeI}.
+To establish the bound on Type II error, we must lower bound $f_0^{\top} \wh{S}^2 f_0$. We first note that by assumption \eqref{eqn:ls_fixed_graph_testing_critical_radius},
 \begin{align*}
@@
 \end{equation*} 
-The upper bound on Type II error~\eqref{eqn:ls_fixed_graph_testing_typeII} then follows from Lemma~\ref{lem:linear_smoother_fixed_graph_testing}.
+The upper bound on Type II error \eqref{eqn:ls_fixed_graph_testing_typeII} then follows from Lemma~\ref{lem:linear_smoother_fixed_graph_testing}.
 
@@
 \label{sec:graph_sobolev_seminorm}
+
 In this section, we prove Lemma~\ref{lem:graph_sobolev_seminorm}, which states an upper bound on $f^{\top} \Lap_{n,r} f$ that holds when $f$ is bounded in Sobolev norm.
@@
 Throughout this proof, we will assume that $f \in H^1(\Xset)$ has zero-mean, meaning $\int_{\Xset} f(x) \,dx = 0$.
-This is without loss of generality---assuming for the moment that~\eqref{eqn:graph_sobolev_seminorm} holds for zero-mean functions, for any $f \in H^1(\Xset)$, taking $a = \int_{\Xset} f(x) \,dx$ and $g = f - a$, we have that
+This is without loss of generality---assuming for the moment that \eqref{eqn:graph_sobolev_seminorm} holds for zero-mean functions, for any $f \in H^1(\Xset)$, taking $a = \int_{\Xset} f(x) \,dx$ and $g = f - a$, we have that
 \begin{equation*}
@@
 
-Now, take $\Omega$ to be an arbitrary bounded open set such that $B(x,c_0) \subset \Omega$ for all $x \in \mc{X}$.
+Now, take $\Omega$ to be an arbitrary bounded open set such that $B(x,c_0) \subseteq \Omega$ for all $x \in \mc{X}$.
 For the remainder of this proof, we will assume that (i) $f \in H^1(\Omega)$ and additionally (ii) $\|f\|_{H^1(\Omega)} \leq C_5 \|f\|_{H^1(\mc{X})}$ for a constant $C_5$ that does not depend on $f$.
@@
 Additionally, we will assume $f \in C^{\infty}(\Omega)$.
-Again, this is without loss of generality, as $C^{\infty}(\Omega)$ is dense in $H^1(\Omega)$ and the expectation on the right hand side of~\eqref{pf:first_order_graph_sobolev_seminorm_1} is continuous in $H^1(\Omega)$.
+Again, this is without loss of generality, as $C^{\infty}(\Omega)$ is dense in $H^1(\Omega)$ and the expectation on the right hand side of \eqref{pf:first_order_graph_sobolev_seminorm_1} is continuous in $H^1(\Omega)$.
 The reason for dealing with a smooth extension $f \in C^{\infty}(\Omega)$ is so that we can make sense of the following equality for any $x$ and $x'$ in $\mc{X}$:
@@
 \end{equation}
-so that it remains now to bound the double integral. Replacing difference by integrated derivative as in~\eqref{pf:first_order_graph_sobolev_seminorm_1.5}, we obtain
+so that it remains now to bound the double integral. Replacing difference by integrated derivative as in \eqref{pf:first_order_graph_sobolev_seminorm_1.5}, we obtain
 \begin{align}
@@
 \end{align*}
-where the last equality follows from the rotational symmetry of $K(\|z\|)$. Plugging back into~\eqref{pf:first_order_graph_sobolev_seminorm_3}, we obtain
+where the last equality follows from the rotational symmetry of $K(\|z\|)$. Plugging back into \eqref{pf:first_order_graph_sobolev_seminorm_3}, we obtain
 \begin{equation*}
@@
 \subsection{Stronger bounds under Lipschitz assumption}
+
 Suppose $f$ satisfies $|f(x') - f(x)| \leq M \|x - x\|$ for all $x,x' \in \mc{X}$.
@@
 \end{proposition}
+
 \paragraph{Proof of Proposition~\ref{prop:graph_sobolev_seminorm_lipschitz}.}
@@
 However, the Sobolev assumption---which gives us control only over $\|\nabla f\|_{\Leb^2(\Xset)}$---will not do the job.
+
 \section{Bounds on neighborhood graph eigenvalues}
 \label{sec:graph_eigenvalues}
+
 In this section, we prove Lemma~\ref{lem:neighborhood_eigenvalue}, following the lead of~\citet{burago2014,trillos2019,calder2019}, who establish similar results with respect to a manifold without boundary.
@@
 See Lemma 28 of~\citet{dunlop2020} for a proof that~\ref{asmp:domain} and~\ref{asmp:density} imply Weyl's Law.
+
 \paragraph{Proof of Lemma~\ref{lem:neighborhood_eigenvalue}.}
@@
 \end{align*}
-Rearranging the above inequality shows that condition~\eqref{eqn:neighborhood_eigenvalue_1} is satisfied.
-
-It is therefore the case that the inequalities in~\eqref{eqn:eigenvalue_bound} hold with probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{m})$.
+Rearranging the above inequality shows that condition \eqref{eqn:neighborhood_eigenvalue_1} is satisfied.
+
+It is therefore the case that the inequalities in \eqref{eqn:eigenvalue_bound} hold with probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{m})$.
 Together, \eqref{eqn:eigenvalue_bound} and \eqref{eqn:weyls_law} imply the following bounds on the graph Laplacian eigenvalues:
@@
 For the rest of this section, we let $\wt{P}_n$ be a probability measure with density $\wt{p}_n$, that satisfies the conclusions of Proposition~\ref{prop:optimal_transport}.
-Additionally we denote by $\wt{T}_n$ an \emph{optimal transport map} between $\wt{P}_n$ and $P_n$, meaning a transportation map which achieves the infimum in~\eqref{eqn:optimal_transport}.
+Additionally we denote by $\wt{T}_n$ an \emph{optimal transport map} between $\wt{P}_n$ and $P_n$, meaning a transportation map which achieves the infimum in \eqref{eqn:optimal_transport}.
 Finally, we write $U_1,\ldots,U_n$ for the preimages of $X_1,\ldots,X_n$ under $\wt{T}_n$, meaning $U_i = \wt{T}_n^{-1}(X_i)$.
@@
 \paragraph{Proof of Theorem~\ref{thm:neighborhood_eigenvalue}.}
-Throughout this proof, we assume that inequalities~\eqref{eqn:dirichlet_energies_1}-\eqref{eqn:isometry_2} are satisfied. We take $A$ and $a$ to be positive constants such that
+Throughout this proof, we assume that inequalities \eqref{eqn:dirichlet_energies_1}-\eqref{eqn:isometry_2} are satisfied. We take $A$ and $a$ to be positive constants such that
 \begin{align*}
@@
 Let $k$ be any number in $1,\ldots,\ell$.
-We start with the upper bound in~\eqref{eqn:eigenvalue_bound}, proceeding as in Proposition 4.4 of \citet{burago2014}.
+We start with the upper bound in \eqref{eqn:eigenvalue_bound}, proceeding as in Proposition 4.4 of \citet{burago2014}.
 Let $f_1,\ldots,f_{k}$ denote the first $k$ eigenfunctions of $\Delta_P$ and set $W := \mathrm{span}\{f_1,\ldots,f_k\}$, so that by the Courant-Fischer principle $D_2(f) \leq \lambda_k(\Delta_P) \|f\|_{P}^2$ for every $f \in W$.
@@
 \end{equation*}
-where the second inequality follows by assumption~\eqref{eqn:neighborhood_eigenvalue_1}.
+where the second inequality follows by assumption \eqref{eqn:neighborhood_eigenvalue_1}.
 
@@
 \end{align*}
-establishing the lower bound in~\eqref{eqn:eigenvalue_bound}.
+establishing the lower bound in \eqref{eqn:eigenvalue_bound}.
 
@@
 \end{align*}
-where the second to last inequality follows from the lower bound $a \lambda_k(G_{n,r}) \leq nr^{d + 2}\lambda_k(\Delta_P)$ that we just derived, and the last inequality from assumption~\eqref{eqn:neighborhood_eigenvalue_1}.
+where the second to last inequality follows from the lower bound $a \lambda_k(G_{n,r}) \leq nr^{d + 2}\lambda_k(\Delta_P)$ that we just derived, and the last inequality from assumption \eqref{eqn:neighborhood_eigenvalue_1}.
 
@@
 \end{align*}
-establishing the upper bound in~\eqref{eqn:eigenvalue_bound}.
+establishing the upper bound in \eqref{eqn:eigenvalue_bound}.
 
@@
 \end{equation*}
-Moreover, ignoring those cells for which $N(Q) = 0$ (since $\wt{P}_n(Q) = 0$ for such $Q$, and so they do not contribute to the essential supremum in~\eqref{eqn:optimal_transport}), appropriately dividing each remaining cell $Q \in \mc{Q}$ into $N(Q)$ subsets $S_1,\ldots,S_{N(Q)}$ of equal volume, and mapping each $S_{\ell}$ to a different design point $X_i \in Q$, we can exhibit a transport map $T$ from $\wt{P}_n$ to $P_n$ for which
+Moreover, ignoring those cells for which $N(Q) = 0$ (since $\wt{P}_n(Q) = 0$ for such $Q$, and so they do not contribute to the essential supremum in \eqref{eqn:optimal_transport}), appropriately dividing each remaining cell $Q \in \mc{Q}$ into $N(Q)$ subsets $S_1,\ldots,S_{N(Q)}$ of equal volume, and mapping each $S_{\ell}$ to a different design point $X_i \in Q$, we can exhibit a transport map $T$ from $\wt{P}_n$ to $P_n$ for which
 \begin{equation*}
@@
 \end{align*}
-Noting that by assumption $P(\wt{Q}) \leq p_{\max} \vol(\wt{Q})$ and $\wt{\delta}^{-d} \leq n$, the claim follows upon plugging back into~\eqref{pf:prop_optimal_transport_3}, and setting
+Noting that by assumption $P(\wt{Q}) \leq p_{\max} \vol(\wt{Q})$ and $\wt{\delta}^{-d} \leq n$, the claim follows upon plugging back into \eqref{pf:prop_optimal_transport_3}, and setting
 \begin{equation*}
@@
 \label{subsec:integrals}
+
 Let us start by making the following observation, which we make use of repeatedly in this section.
@@
 \paragraph{Proof of Lemma~\ref{lem:tau_bound}.}
-We first establish our estimates of $\tau_h(x)$, and then upper bound $\|\nabla\tau_h(x)\|$. Using~\eqref{eqn:integral_boundary}, we have that
+We first establish our estimates of $\tau_h(x)$, and then upper bound $\|\nabla\tau_h(x)\|$. Using \eqref{eqn:integral_boundary}, we have that
 \begin{align*}
@@
 
-To establish~\eqref{eqn:smoothening_error_energy}, noting that $\Lambda_ha = a$ for any $a \in \Reals$, we have that
+To establish \eqref{eqn:smoothening_error_energy}, noting that $\Lambda_ha = a$ for any $a \in \Reals$, we have that
 \begin{align*}
@@
 \end{equation*}
-Then integrating over $\Xset$ with respect to $p$ yields~\eqref{eqn:smoothening_error_energy}.
+Then integrating over $\Xset$ with respect to $p$ yields \eqref{eqn:smoothening_error_energy}.
 
@@
 \label{subsec:random_functionals}
+
 We will use Lemma~\ref{lem:poincare} in the proof of Proposition~\ref{prop:isometry}. 
@@
 	\label{lem:poincare}
-	Let $U \subset \Xset$ be a measurable subset such that $\mathrm{vol}(U) > 0$, and $\diam(U) \leq 2A_0\wt{\delta}$.
+	Let $U \subseteq \Xset$ be a measurable subset such that $\mathrm{vol}(U) > 0$, and $\diam(U) \leq 2A_0\wt{\delta}$.
 Then, letting $a = (\wt{P}_n(U))^{-1} \cdot \int_{U} f(x) \wt{p}_n(x) \,dx$ be the average of $f$ over $U$, it holds that
@@
 Now, since $x'$ and $x$ belong to $U$, we have that $\|x' - x\| \leq 2A_0\wt{\delta}$.
-Set $V = B(x,r) \cap B(x',r)$, and note that $B(x,r - 2A_0\wt{\delta}) \subset V$.
+Set $V = B(x,r) \cap B(x',r)$, and note that $B(x,r - 2A_0\wt{\delta}) \subseteq V$.
 Moreover, $r - 2A_0\wt{\delta} \leq r \leq c_0$ by assumption.
-Therefore by~\eqref{eqn:integral_boundary},
+Therefore by \eqref{eqn:integral_boundary},
 \begin{equation*}
@@
 
-Plugging the upper bound~\eqref{pf:poincare_2} back into~\eqref{pf:poincare_1}, we have that
+Plugging the upper bound \eqref{pf:poincare_2} back into \eqref{pf:poincare_1}, we have that
 \begin{align*}
@@
 
-\subsection{Proof of Propositions~\ref{prop:dirichlet_energies} and~\ref{prop:isometry}}
+\subsection{Proof of Propositions~\ref{prop:dirichlet_energies} and \ref{prop:isometry}}
 \label{subsec:proof_of_prop_dirichlet_energies_and_isometry}
@@
 \end{align*}
-where $(i)$ follows from Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, $(ii)$ follows from~\eqref{eqn:calder19_1}, and $(iii)$ follows from~Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}.
+where $(i)$ follows from Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, $(ii)$ follows from \eqref{eqn:calder19_1}, and $(iii)$ follows from~Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}.
 
@@
 \end{align*}
-where $(iv)$ follows from Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}, $(v)$ follows from~\eqref{eqn:calder19_1}, and $(vi)$ follows from the proof of Lemma~\ref{lem:graph_sobolev_seminorm}.
+where $(iv)$ follows from Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}, $(v)$ follows from \eqref{eqn:calder19_1}, and $(vi)$ follows from the proof of Lemma~\ref{lem:graph_sobolev_seminorm}.
 
 \paragraph{Proof of Proposition~\ref{prop:isometry}.}
-\mbox{}\\
-\mbox{}\\
 \textit{Proof of (1).}
-We begin by upper bounding $\bigl\|\wt{P}f\bigr\|_{n}$. By the Cauchy-Schwarz inequality and the bound on $\|\wt{p}_n - p\|_{\infty}$ in~\eqref{eqn:optimal_transport_2},
+We begin by upper bounding $\bigl\|\wt{P}f\bigr\|_{n}$. By the Cauchy-Schwarz inequality and the bound on $\|\wt{p}_n - p\|_{\infty}$ in \eqref{eqn:optimal_transport_2},
 \begin{align*}
@@
 \end{equation}
-Now, noticing that $\bigl\|\wt{P}f\bigr\|_{n} = \bigl\|\wt{P}^{\star}\wt{P}f\bigr\|_{\wt{P}_n}$, we can use the upper bound~\eqref{pf:prop_isometry_1} to show that
+Now, noticing that $\bigl\|\wt{P}f\bigr\|_{n} = \bigl\|\wt{P}^{\star}\wt{P}f\bigr\|_{\wt{P}_n}$, we can use the upper bound \eqref{pf:prop_isometry_1} to show that
 \begin{align}
@@
 \end{align}
-where $(i)$ follows from~\eqref{eqn:calder19_2} and $(ii)$ follows from~\eqref{eqn:calder19_2} and ~\eqref{pf:prop_isometry_1}.
+where $(i)$ follows from \eqref{eqn:calder19_2} and $(ii)$ follows from \eqref{eqn:calder19_2} and  \eqref{pf:prop_isometry_1}.
 
@@
 where the latter inequality follows from the proof of Proposition~\ref{prop:dirichlet_energies}, Part (2).
-Then Proposition~\ref{prop:isometry}, Part (1) follows by plugging this inequality into~\eqref{pf:prop_isometry_2} and taking
+Then Proposition~\ref{prop:isometry}, Part (1) follows by plugging this inequality into \eqref{pf:prop_isometry_2} and taking
 \begin{equation*}
@@
 \textit{Proof of (2).}
-By the triangle inequality and~\eqref{eqn:calder19_2},
+By the triangle inequality and \eqref{eqn:calder19_2},
 \begin{align}
@@
 \end{align}
-where $(iii)$ follows by the triangle inequality, $(iv)$ follows from Lemma~\ref{lem:smoothening_error}, and $(v)$ follows from~\eqref{eqn:calder19_1} and Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}.
-On the other hand, by~\eqref{eqn:calder19_2} and Lemma~\ref{lem:smoothening_error},
+where $(iii)$ follows by the triangle inequality, $(iv)$ follows from Lemma~\ref{lem:smoothening_error}, and $(v)$ follows from \eqref{eqn:calder19_1} and Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}.
+On the other hand, by \eqref{eqn:calder19_2} and Lemma~\ref{lem:smoothening_error},
 \begin{align*}
@@
 \end{align*}
-and plugging this estimate along with~\eqref{pf:prop_isometry_5} back into~\eqref{pf:prop_isometry_4}, upon choosing
+and plugging this estimate along with \eqref{pf:prop_isometry_5} back into \eqref{pf:prop_isometry_4}, upon choosing
 \begin{equation*}
@@
 \paragraph{Proof of Lemma~\ref{lem:empirical_norm_sobolev}.}
-In this proof, we will find it more convenient to deal with the parameterization $b = 1/\delta$. To establish~\eqref{eqn:empirical_norm_sobolev}, it is sufficient to show that
+In this proof, we will find it more convenient to deal with the parameterization $b = 1/\delta$. To establish \eqref{eqn:empirical_norm_sobolev}, it is sufficient to show that
 \begin{equation*}
@@
 
-Now, suppose $f \in \Leb^4(\Xset)$ satisfies~\eqref{pf:empirical_norm_sobolev_1} only when $t = 1$.
-For each $k = 1,2,\ldots$ let $f_k := n^{1/(2k)}f$, so that each $f_k$ satisfies~\eqref{pf:empirical_norm_sobolev_1} with respect to $t = 1 - 1/k$.
+Now, suppose $f \in \Leb^4(\Xset)$ satisfies \eqref{pf:empirical_norm_sobolev_1} only when $t = 1$.
+For each $k = 1,2,\ldots$ let $f_k := n^{1/(2k)}f$, so that each $f_k$ satisfies \eqref{pf:empirical_norm_sobolev_1} with respect to $t = 1 - 1/k$.
 Clearly $\|f_k - f\|_{\Leb^4(\Xset)} \to 0$ as $k \to \infty$, and therefore
@@
 \label{sec:manifold}
+
 In this section, we restate a few results of \citet{trillos2019,calder2019}, which are analogous to Lemmas~\ref{lem:graph_sobolev_seminorm} and~\ref{lem:neighborhood_eigenvalue} but cover the case where $\Xset$ is an $m$-dimensional submanifold without boundary.
@@
 
-\section{Proofs of Main Results}
+\section{Proofs of main results}
 \label{sec:main_results}
+
 We are now in a position to prove Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing_manifold}, as well as a few other claims from our main text.
@@
 \paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}.}
-We have shown that the inequalities~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} are satisfied with probability at least $1 - \delta - C_1n\exp(-c_1nr^d)$, and throughout this proof we take as granted that both of these inequalities hold.
+We have shown that the inequalities \eqref{eqn:graph_sobolev_seminorm} and \eqref{eqn:neighborhood_eigenvalue} are satisfied with probability at least $1 - \delta - C_1n\exp(-c_1nr^d)$, and throughout this proof we take as granted that both of these inequalities hold.
 
 Now, set $\wt{\rho} = M^{-4/(2 + d)}n^{-2/(2 + d)}$ as prescribed in Theorem~\ref{thm:laplacian_smoothing_estimation1}, and note that $\wt{\rho}^{-d/2} \leq n$ is implied by the assumption $M \leq n^{1/d}$.
-Therefore  from~\eqref{eqn:neighborhood_eigenvalue} and Lemma~\ref{lem:variance_term_estimation}, it follows that
+Therefore  from \eqref{eqn:neighborhood_eigenvalue} and Lemma~\ref{lem:variance_term_estimation}, it follows that
 \begin{equation*}
@@
 \end{equation*}
-As a result, by Lemma~\ref{lem:ls_fixed_graph_estimation} along with~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue}, with probability at least $1 - \delta - C_1n\exp(-c_1nr^d) - \exp(-\wt{\rho}^{-d/2}/8C_3^2)$ it holds that,
+As a result, by Lemma~\ref{lem:ls_fixed_graph_estimation} along with \eqref{eqn:graph_sobolev_seminorm} and \eqref{eqn:neighborhood_eigenvalue}, with probability at least $1 - \delta - C_1n\exp(-c_1nr^d) - \exp(-\wt{\rho}^{-d/2}/8C_3^2)$ it holds that,
 \begin{align}
@@
 \end{align}
-The first term on the right hand side of~\eqref{pf:laplacian_smoothing_estimation1_1} is a bias term, while the second, third, and fourth terms each contribute to the variance.
+The first term on the right hand side of \eqref{pf:laplacian_smoothing_estimation1_1} is a bias term, while the second, third, and fourth terms each contribute to the variance.
 Of these, under our assumptions the third term dominates, as we show momentarily.
@@
 \end{equation*}
-Then plugging this upper bound back into~\eqref{pf:laplacian_smoothing_estimation1_1}, we have that
+Then plugging this upper bound back into \eqref{pf:laplacian_smoothing_estimation1_1}, we have that
 \begin{align*}
@@
 
-Continuing on from~\eqref{pf:laplacian_smoothing_estimation1_1}, from Lemma~\ref{lem:variance_term_estimation} we have that
+Continuing on from \eqref{pf:laplacian_smoothing_estimation1_1}, from Lemma~\ref{lem:variance_term_estimation} we have that
 \begin{equation*}
@@
 
-Once again continuing on from~\eqref{pf:laplacian_smoothing_estimation1_1}, from Lemma~\ref{lem:variance_term_estimation} we have that,
+Once again continuing on from \eqref{pf:laplacian_smoothing_estimation1_1}, from Lemma~\ref{lem:variance_term_estimation} we have that,
 \begin{equation*}
@@
 Therefore Proposition~\ref{prop:out_of_sample_error} together with Theorem~\ref{thm:laplacian_smoothing_estimation1} implies that with high probability, $\wh{f}$ achieves the nearly-optimal (up to a factor of $\log n$) estimation rates out-of-sample error---that is, $\|\wh{f} - f_0\|_{\Leb^2(\Xset)}^2 \leq C \log(n) M^{2d/(2 + d)}n^{-2/(2+d)}$---as long as $M \leq Cn^{1/d}$.
+
 \paragraph{Proof of Proposition~\ref{prop:out_of_sample_error}.}
@@
 where $C_{o}$ is a constant given in Lemma~16 of \citet{chaudhuri2010}.
-Note that for $n$ sufficiently large, $\varepsilon_n \leq c_0$, and therefore by~\eqref{eqn:integral_boundary} we have that for every $x \in \Xset$, $P(B(x,\varepsilon_n)) \geq 2C_{o}\log(1/\delta)d\frac{\log n}{n}$.
+Note that for $n$ sufficiently large, $\varepsilon_n \leq c_0$, and therefore by \eqref{eqn:integral_boundary} we have that for every $x \in \Xset$, $P(B(x,\varepsilon_n)) \geq 2C_{o}\log(1/\delta)d\frac{\log n}{n}$.
 Consequently, by Lemma~16 of \citet{chaudhuri2010} it holds that with probability at least $1 - \delta$,
@@
 \end{equation}
-But if~\eqref{pf:out_of_sample_error_1} is true, it must also be true that for each $i = 1,\ldots,n$ and for every $x \in V_i$, the distance $\|x - X_i\| \leq \varepsilon_n$.
+But if \eqref{pf:out_of_sample_error_1} is true, it must also be true that for each $i = 1,\ldots,n$ and for every $x \in V_i$, the distance $\|x - X_i\| \leq \varepsilon_n$.
 Thus by the triangle inequality, $\max_{i = 1,\ldots,n} \mathrm{diam}(V_i) \leq 2\varepsilon_n$.
-Plugging back in to~\eqref{pf:out_of_sample_error_0}, and using the upper bound volume $\vol(V_i) \leq \nu_d \bigl(\mathrm{diam}(V_i)\bigr)^d$, we obtain the desired upper bound on $\|\wh{f} - f\|_{\Leb^2(\Xset)}^2$.
+Plugging back in to \eqref{pf:out_of_sample_error_0}, and using the upper bound volume $\vol(V_i) \leq \nu_d \bigl(\mathrm{diam}(V_i)\bigr)^d$, we obtain the desired upper bound on $\|\wh{f} - f\|_{\Leb^2(\Xset)}^2$.
 
 \paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_estimation_manifold}.}
-The proof of Theorem~\ref{thm:laplacian_smoothing_estimation_manifold} follows exactly the same steps as the proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}, replacing the references to Lemma~\ref{lem:graph_sobolev_seminorm} and~\ref{lem:neighborhood_eigenvalue} by references to Proposition~\ref{prop:garciatrillos19_1} and~\eqref{eqn:neighborhood_eigenvalue_manifold}, and the ambient dimension $d$ by the intrinsic dimension $m$.
+The proof of Theorem~\ref{thm:laplacian_smoothing_estimation_manifold} follows exactly the same steps as the proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}, replacing the references to Lemma~\ref{lem:graph_sobolev_seminorm} and~\ref{lem:neighborhood_eigenvalue} by references to Proposition~\ref{prop:garciatrillos19_1} and \eqref{eqn:neighborhood_eigenvalue_manifold}, and the ambient dimension $d$ by the intrinsic dimension $m$.
 
@@
 Let $\delta = 1/b$.
-Recall that we have shown that the inequalities~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} are satisfied with probability at least $1 - 1/b - C_1n\exp(-c_1nr^d)$, and throughout this proof we take as granted that both of these inequalities hold.
-
-Now, we would like to invoke Lemma~\ref{lem:ls_fixed_graph_testing}, and in order to do so, we must show that the inequality~\eqref{eqn:ls_fixed_graph_testing_critical_radius} is satisfied with respect to $G = G_{n,r}$.
+Recall that we have shown that the inequalities \eqref{eqn:graph_sobolev_seminorm} and \eqref{eqn:neighborhood_eigenvalue} are satisfied with probability at least $1 - 1/b - C_1n\exp(-c_1nr^d)$, and throughout this proof we take as granted that both of these inequalities hold.
+
+Now, we would like to invoke Lemma~\ref{lem:ls_fixed_graph_testing}, and in order to do so, we must show that the inequality \eqref{eqn:ls_fixed_graph_testing_critical_radius} is satisfied with respect to $G = G_{n,r}$.
 First, we upper bound the right hand side of this inequality.
-Set $\wt{\rho} = M^{-8/(4 + d)}n^{-4/(4 + d)}$ as prescribed by Theorem~\ref{thm:laplacian_smoothing_testing}, it follows from~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} that
+Set $\wt{\rho} = M^{-8/(4 + d)}n^{-4/(4 + d)}$ as prescribed by Theorem~\ref{thm:laplacian_smoothing_testing}, it follows from \eqref{eqn:graph_sobolev_seminorm} and \eqref{eqn:neighborhood_eigenvalue} that
 \begin{align*}
@@
 
-Next we lower bound the left hand side of the inequality~\eqref{eqn:ls_fixed_graph_testing_critical_radius}---i.e.
+Next we lower bound the left hand side of the inequality \eqref{eqn:ls_fixed_graph_testing_critical_radius}---i.e.
 we lower bound the empirical norm $\|f_0\|_n^2$---using Lemma~\ref{lem:empirical_norm_sobolev}.
 Recall that by assumption, $M \leq M_{\max}(d)$.
-Therefore, taking $C \geq C_6$ in~\eqref{eqn:laplacian_smoothing_testing}, implies that the lower bound on $\|f\|_{\Leb^2(\mc{X})}$ in~\eqref{eqn:empirical_norm_sobolev_1} is satisfied.
-As a result, it follows from~\eqref{eqn:empirical_norm_sobolev} that
+Therefore, taking $C \geq C_6$ in \eqref{eqn:laplacian_smoothing_testing}, implies that the lower bound on $\|f\|_{\Leb^2(\mc{X})}$ in \eqref{eqn:empirical_norm_sobolev_1} is satisfied.
+As a result, it follows from \eqref{eqn:empirical_norm_sobolev} that
 \begin{equation*}
@@
 with probability at least $1 - 5/b$.
-Taking $C \geq C_2 + 4 + (4\sqrt{2})/c_3^{2} + 4/c_3^2$ in~\eqref{eqn:laplacian_smoothing_testing} thus implies~\eqref{eqn:ls_fixed_graph_testing_critical_radius}, and we may therefore use Lemma~\ref{lem:ls_fixed_graph_testing} to upper bound the type II error the Laplacian smoothing test $\wh{\varphi}$.
-Observe that by~\eqref{eqn:neighborhood_eigenvalue} and the lower bound in Lemma~\ref{lem:variance_term_testing},
+Taking $C \geq C_2 + 4 + (4\sqrt{2})/c_3^{2} + 4/c_3^2$ in \eqref{eqn:laplacian_smoothing_testing} thus implies \eqref{eqn:ls_fixed_graph_testing_critical_radius}, and we may therefore use Lemma~\ref{lem:ls_fixed_graph_testing} to upper bound the type II error the Laplacian smoothing test $\wh{\varphi}$.
+Observe that by \eqref{eqn:neighborhood_eigenvalue} and the lower bound in Lemma~\ref{lem:variance_term_testing},
 \begin{equation*}
@@
 \paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_testing_manifold}.}
-The proof of Theorem~\ref{thm:laplacian_smoothing_testing_manifold} follows exactly the same steps as the proof of Theorem~\ref{thm:laplacian_smoothing_testing}, replacing the references to Lemma~\ref{lem:graph_sobolev_seminorm} and~\ref{lem:neighborhood_eigenvalue} by references to Propositions~\ref{prop:garciatrillos19_1} and~\eqref{eqn:neighborhood_eigenvalue_manifold}, and the ambient dimension $d$ by the intrinsic dimension $m$.
-
-\paragraph{Proof of~\eqref{eqn:laplacian_smoothing_testing_low_smoothness}.}
+The proof of Theorem~\ref{thm:laplacian_smoothing_testing_manifold} follows exactly the same steps as the proof of Theorem~\ref{thm:laplacian_smoothing_testing}, replacing the references to Lemma~\ref{lem:graph_sobolev_seminorm} and~\ref{lem:neighborhood_eigenvalue} by references to Propositions~\ref{prop:garciatrillos19_1} and \eqref{eqn:neighborhood_eigenvalue_manifold}, and the ambient dimension $d$ by the intrinsic dimension $m$.
+
+\paragraph{Proof of \eqref{eqn:laplacian_smoothing_testing_low_smoothness}.}
 When $\rho = 0$, the Laplacian smoother $\wh{f} = \mathbf{Y}$, the test statistic $\wh{T} = \frac{1}{n}\|\mathbf{Y}\|_2^2$, and the threshold $\wh{t}_b = 1 + 2bn^{-1/2}$.
@@
 \label{subsec:convenient_estimate}
+
 The following Lemmas provides convenient upper and lower bounds on our estimation variance term (Lemma~\ref{lem:variance_term_estimation}) and testing variance term (Lemma~\ref{lem:variance_term_testing}).
@@
 \end{lemma}
+
 \paragraph{Proof of Lemma~\ref{lem:variance_term_estimation}.}
@@
 
-\section{Concentration Inequalities}
+\section{Concentration inequalities}
+
 \label{sec:concentration}
@@
 	\end{equation}
-	Applying~\eqref{eqn:paley_zygmund_pf1} with $Z = \norm{f}_n^2$, $p = 1$, $q = 2$ and $\lambda = \frac{1}{b}$, by assumption~\eqref{eqn:paley_zygmund_1} we have
+	Applying \eqref{eqn:paley_zygmund_pf1} with $Z = \norm{f}_n^2$, $p = 1$, $q = 2$ and $\lambda = \frac{1}{b}$, by assumption \eqref{eqn:paley_zygmund_1} we have
 	\begin{equation*}
@@
 \end{lemma}
+
 \paragraph{Proof of Lemma~\ref{lem:max_degree}.}
@@
 \end{align*}
-with the second inequality following from~\eqref{eqn:integral_boundary}, and the final inequality from the normalization $\int_{\Rd} K(\|z\|) \,dz = 1$. Similar derivations yield the upper bound
+with the second inequality following from \eqref{eqn:integral_boundary}, and the final inequality from the normalization $\int_{\Rd} K(\|z\|) \,dz = 1$. Similar derivations yield the upper bound
 \begin{equation*}
@@
 \end{equation*} 
-and plugging these bounds in to~\eqref{pf:max_degree_1}, we determine that
+and plugging these bounds in to \eqref{pf:max_degree_1}, we determine that
 \begin{equation*}

### 2021-02-28 manuscripts/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
Typos and copyedits

@@
 In this paper, we will adopt the random design perspective, and seek to answer the following question:
-\begin{quote}{When $f_0$ is assumed to be smooth in a continuum sense, does Laplacian smoothing achieve optimal performance for estimation and goodness-of-fit testing?} 
+\begin{quote}{When we assume the regression function $f_0$ to be smooth in a continuum sense, does Laplacian smoothing achieve optimal performance for estimation and goodness-of-fit testing?} 
 \end{quote}
@@
 This will clearly not lead to a consistent estimator of $f_0$ across the design points (as it always yields $Y_i$ at each $X_i$).
-In this light, our results for $d>1$ favorably distinguish Laplacian smoothing from its natural variational analog.
+In this light, our results when $d>1$ favorably distinguish Laplacian smoothing from its natural variational analog.
 
@@
 We leave analysis of the Sobolev case to future work.
-	\item Under the assumption that $f_0$ is Lipschitz smooth, we can also replace the factor of $\delta$ in the high probability bound by the sharper $\delta^2/n$.
+	\item When $f_0$ is Lipschitz smooth, we can also replace the factor of $\delta$ in the high probability bound by the sharper $\delta^2/n$.
 \end{itemize}
@@
 
-Of course, we know this is not the case: $\wh{f}$ and $\wt{f}$ look very different when $d > 1$.
+Of course, we know this is not the case: \smash{$\wh{f}$} and \smash{$\wt{f}$} look very different when $d > 1$.
 What is driving this difference?

### 2021-03-01 manuscripts/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
Fix typo, bib copyedit

@@
 	\label{lem:neighborhood_eigenvalue} 
-  There are constants $N,C_1,C_3,c_1,c_3 > 0$ such that for $n \geq N$, $C_0(\log n/n)^{1/d} \leq r \leq c_0$, and $f \in H^1(\Xset)$, with probability at least $1 - C_1n\exp(-c_1nr^d)$, it holds that
+  There are constants $N,C_1,C_3,c_1,c_3 > 0$ such that for $n \geq N$ and $C_0(\log n/n)^{1/d} \leq r \leq c_0$, with probability at least $1 - C_1n\exp(-c_1nr^d)$, it holds that
 	\begin{equation}

### 2021-03-02 manuscripts/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
Typography
Funding acknowledgement
Remove refs to continuity

@@
 \usepackage{natbib}
-\renewcommand{\bibname}{REFERENCES}
+\renewcommand{\bibname}{References}
 \renewcommand{\bibsection}{\subsubsection*{\bibname}}
@@
 \end{equation}
-where the optimization is performed over all functions for which the criterion is well-defined and finite (continuous functions $f$ that have a weak derivative $\nabla f$ in $L^2(\Xset)$).
+where the optimization is performed over all functions for which the criterion is well-defined and finite (functions $f$ whose weak derivative $\nabla f$ in $L^2(\Xset)$).
 Analogously, for testing, we could use:
@@
 \end{equation*}
-We now state Theorem~\ref{thm:laplacian_smoothing_testing}, our main testing result.
+
+We now give Theorem~\ref{thm:laplacian_smoothing_testing}, our main testing result.
 \begin{theorem}
@@
 \end{equation*}
-and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies e.g., Dirichlet boundary conditions.} To be clear, this argument does not formally imply that $\wh{f}$ and $\wt{f}$ are close (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of continuous functions with weak derivatives in $\Leb^2(\Xset)$).
-But it does seem to suggest that the two estimators should behave somewhat similarly.
+and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies e.g., Dirichlet boundary conditions.} To be clear, this argument does not formally imply that $\wh{f}$ and $\wt{f}$ are close (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of functions with weak derivatives in $\Leb^2(\Xset)$).
+However, it does seem to suggest that the two estimators should behave somewhat similarly.
 
@@
 
+\subsection*{Acknowledgements}
+
+AG and RJT were supported by ONR grant N00014-20-1-2787.
+
 \bibliographystyle{plainnat}

### 2021-03-02 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
More progress made on the draft of Laplacian Eigenmaps paper.

@@
 \begin{equation*}
-H^s(P;M) := \{f \in H^s(P): \|f\|_{H^s(P)} \leq M\}
+H^s(P;M) := \bigl\{f \in H^s(P): \|f\|_{H^s(P)} \leq M\bigr\}
 \end{equation*}
@@
 	A test constructed using $\wh{T}$ has non-trivial power whenever $f_0 \in H^1(\Xset;M)$ satisfies $\norm{f_0}_{P}^2 \gtrsim M^{2d/(2 + d)}n^{-4/(4 + d)}$ and $d < 4$.
-	\item \textbf{Higher-order smoothness conditions.} When the regression function $f_0 \in H_0^s(\Xset;M)$ and additionally the density $p \in C^{r}(\mc{X};M)$, with high probability the error $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim M^{2d/(2s + d)}n^{-2s/(2s + d)}$.
+	\item \textbf{Higher-order smoothness conditions.} When the regression function $f_0 \in H_0^s(\Xset;M)$ and additionally the density $p \in C^{r}(\mc{X};M)$, with high probability the error $\norm{\wh{f} - f_0}_{n}^2 \lesssim M^{2d/(2s + d)}n^{-2s/(2s + d)}$.
 Additionally, a test constructed using $\wh{T}$ has non-trivial power whenever $f_0 \in H_0^s(\Xset;M)$ satisfies $\norm{f_0}_{P}^2 \gtrsim M^{2d/(4s + d)}n^{-4s/(4s + d)}$ and $d < 4$.
 	\item \textbf{Manifold adaptivity.}
-	If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$ and $s \leq 2$, each of the aforementioned rates hold with $d$ replaced by $m$.
+	If $\mc{X} \subset \Rd$ is a manifold of dimension $m < d$, and if $H^s(\mc{X})$ for $s \leq 2$, then each of the aforementioned rates hold with $d$ replaced by $m$.
 \end{itemize}
@@
 In this case the estimator reduces to a classical spectral series estimator, and there is no error incurred by approximating $\Delta_P$.
-\cite{lee2016} considers a similar estimator---using a different normalization of the Laplacian $L$---in both the supervised and semi-supervised setups, but get suboptimal rates for the supervised problem.
-To the best of knowledge, there has been no analysis of the test statistic $\wh{T}$ in the random design framework which we study.
+\cite{lee2016} consider a similar estimator---using eigenvectors of a different normalization of the Laplacian $L$---in both the supervised and semi-supervised setups, but get suboptimal rates for the supervised problem.
+As far as we know, there has been no analysis of the test statistic $\wh{T}$ in the random design framework which we study.
 	\item \textcolor{red}{(TODO)}: Alternatively, \cite{trillos2020} \textcolor{blue}{(Green 2021)} use the graph Laplacian operator to induce a penalty over functions $f: \{\bf X\} \to \Reals$, and study a regularized estimator estimator, which minimizes the penalized loss
@@
 \begin{itemize}
-	\item \textbf{(1.1) How to tune hyper-parameters.} We note that the ranges~\eqref{eqn:radius_fo} and~\eqref{eqn:eigenvector_estimation_fo} depend on unknown quantities such as $f_0$ and $P$.
+	\item Regarding the restrictions on hyperparameters: the lower bound on $\varepsilon$ is the connectivity threshold, the smallest length scale at which the resulting graph will still, with high probability, be connected.
+On the other hand, as we will see in Section~\ref{sec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K \asymp \lambda_K(\Delta_P) \asymp K^{2/d}$.
+Finally, the restriction $K \asymp (M^2n)^{2d/(2 + d)}$ is chosen  to optimally trade-off bias and variance.
+Together, these restrictions suggest that the
+	\item \textbf{(1.1) How to tune hyper-parameters.} We note that the ranges~\eqref{eqn:radius_fo} and~\eqref{eqn:eigenvector_estimation_fo} depend on unknown quantities such as the dimension $d$ and radius of the Sobolev ball $M$.
 In practice, one typically tunes hyper-parameters by sample-splitting or cross-validation.
 However, because the estimator $\wh{f}$ is defined only in-sample, such approaches cannot be straightforwardly applied to select the graph radius $\varepsilon$, or number of eigenvectors $K$.
-We discuss out-of-sample extensions of $\wh{f}$ later, in Section~\ref{sec:out_of_sample}.
+We return to this issue upon considering out-of-sample extensions of $\wh{f}$ later, in Section~\ref{sec:out_of_sample}.
 	\item \textbf{(1.2) The dependence on failure probability can be improved.} The upper bound on in-sample risk given by Equation~\eqref{eqn:laplacian_eigenmaps_estimation_fo} depends on the pre-factor $C/\delta$, and holds with probability $1 - \delta - \textcolor{red}{(?)}$.
@@
 \begin{itemize}
-	\item \textbf{(2.1) When $d \geq 4$, the testing problem is different.} When $d \geq 4$, it is not the case that $\Leb^4(\mc{X})$ continuously embeds into $H^1(\mc{X})$, and testing rates over $H^1(\mc{X})$ are not known.
+	\item \textbf{(2.1) When $d \geq 4$, the testing problem is different.} When $d \geq 4$, it is not the case that $H^1(\mc{X})$ continuously embeds into $\Leb^4(\mc{X})$, and as far we know the optimal testing rates over $H^1(\mc{X})$ are not known.
 On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then simply taking $T = \|Y\|_n^2$ achieves the optimal rates.
@@
 Under appropriate choices of $\varepsilon$ and $K$, the estimator $\wh{f}$ has (in-sample) optimal error (up to constant factors) $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2s/(2s + d)}$.
+
+\begin{enumerate}[label=(A\arabic*)]
+	\setcounter{enumi}{5}
+	\item 
+	The graph radius $\varepsilon$ and number of eigenvectors $K$ satisfies
+	\begin{equation*}
+	C_0 \cdot \max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/d}, n^{-1/(2r + d)}\biggr\} \leq \varepsilon \leq \min\{i_0, K^{-1/d}\}
+	\end{equation*}
+	and
+	\begin{equation*}
+	K \asymp (M^2n)^{2d/(2s + d)}
+	\end{equation*}
+\end{enumerate}
+
 \begin{theorem}
@@
 For instance, it is not a Reproducing Kernel Hilbert Space, nor does it embed into any H\"{o}lder space.
-As a result, one has to be a bit careful.
-Certain versions of the nonparametric regression problem are ill-posed---for instance when loss is measured in $\Leb^{\infty}$ norm, or when the design points $X$ are assumed to be fixed.
-It also takes certain estimators ``off the table'', most notably RKHS-based methods such as smoothing splines.
+As a result, certain versions of the nonparametric regression problem are ill-posed---for instance when loss is measured in $\Leb^{\infty}$ norm, or when the design points $X$ are assumed to be fixed.
+Likewise, certain estimators are ``off the table'', most notably RKHS-based methods such as smoothing splines.
 However, at these scales spectral series estimators still obtain the optimal rates for random design regression measured in $\Leb^2(P)$ norm.
@@
 Both requirements are related and fundamental, as we shall now explain.
-As previously mentioned, $L$ converges to a weighted-version of the Laplace operator $\Delta_P$ (also known as a Fokker-Planck) operator.
+As previously mentioned, $L$ converges to a weighted-version of the Laplace operator $\Delta_P$ (also known as the Fokker-Planck) operator.
 The operator $\Delta_P$ includes both first and second-order derivatives,
@@
 Therefore there are scalings of $s$ and $d$ for which we do not get the optimal rate.
-We do not know whether this due to looseness in our proof techniques, or a problem with the test itself.
+We do not know whether this is due to looseness in our proof techniques, or a problem with the test itself, and leave this for future work.
 
@@
 \textbf{(1) Bias and variance terms, for estimation and testing}.
-We analyze the estimation error of $\wh{f}$ by first condition on the design points $X$ and computing \emph{design-dependent} bias and variance terms.
+We analyze the estimation error of $\wh{f}$, and the testing error of $\wh{\varphi}$, by first conditioning on the design points $X$ and deriving \emph{design-dependent} bias and variance terms.
 For estimation, we have that with high probability, conditional on the design $X$,
@@
 
-The semi-norm $\dotp{L^s f_0}{f_0}_n$ and eigenvalue $\lambda_{K}$ are random variables that depend the design $X$. It remains to establish suitable upper and lower bounds on these quantities.
+The semi-norm $\dotp{L^s f_0}{f_0}_n$ and eigenvalue $\lambda_{K}$ are random variables that depend the random design points $X_1,\ldots,X_n$.
+It remains to establish suitable upper and lower bounds on these quantities.
 
@@
 The proof of Proposition~\ref{prop:graph_seminorm} involves two steps.
-In the first step, we show that with probability at least $1 - \delta$, the graph semi-norm $\dotp{L^s f}{f}_n$ is upper bounded by a non-local continuum energy $E_{\varepsilon}^{(s)}(f)$.
-This non-local continuum energy is defined as the following semi-norm,
+In the first step, we show that with probability at least $1 - \delta$, the graph semi-norm $\dotp{L^s f}{f}_n$ is upper bounded by a non-local continuum energy $E_{\varepsilon}^{(s)}(f)$, defined as
 \begin{equation*}
-E_{\varepsilon}^{(s)}(f) := \dotp{L_{P,\varepsilon}^sf}{f}_{P},
+E_{\varepsilon}^{(s)}(f) := \dotp{L_{P,\varepsilon}^sf}{f}_{P}
 \end{equation*}
@@
 \begin{equation*}
-L_{P,\varepsilon}f(x) := \int \bigl(f(z) - f(x)\bigr) \eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dP(x).
+L_{P,\varepsilon}f(x) := \int_{\mc{X}}\bigl(f(z) - f(x)\bigr) \eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dP(x).
 \end{equation*}
@@
 
-When $s = 1$, these results are somewhat standard~\textcolor{red}{(references)}.
-On the other hand, much less work has been done in the $s > 1$ case, and the analysis is more complicated.
+When $s = 1$, similar results have been shown in \textcolor{red}{(?)}.
+On the other hand, much less work has been done in the $s > 1$ case, where the analysis is more complicated.
 For one thing, we note that when $s > 1$ the graph energy $\dotp{L^s f}{f}_n$ is itself a biased estimate of the non-local energy $E_{\varepsilon}^{(s)}(f)$.
-This bias term is the second term on the right hand side of~\eqref{eqn:graph_seminorm_1}, and is neglible when $r \gtrsim n^{-1/(2r + d)}$.
+This bias term is the second term on the right hand side of~\eqref{eqn:graph_seminorm_1}, and is neglible only when $r \gtrsim n^{-1/(2r + d)}$.
 On the other hand, in order to give an estimate of $E_{\varepsilon}^{(s)}(f)$ in terms of the Sobolev norm $D_s(f)$, we first show that $L_{P,\varepsilon}^jf(x) \approx \Delta_P^jf(x)$, for $j = s/2$ when $s$ is odd and $j = (s - 1)/2$ when $s$ is even.
-In words, we show that a $j$th-order iterated difference operator approximates a $2j$-th order differential operator.
-It is in order to establish this (somewhat remarkable) fact that we require on the one hand $p \in C^{r}(\mc{X};M)$, and on the other hand that $f$ be zero-trace.
+In words, we show that a $j$th-order iterated difference operator approximates a $2j$-th (or $2j + 1$-th) order differential operator.
+It is in order to establish this fact that we require on the one hand $p \in C^{r}(\mc{X};M)$, and on the other hand that $f$ be zero-trace.
 
@@
 \subsection{Riemmanian manifolds}
-We assume $\mc{X}$ is a closed, connected, smooth manifold without boundary of dimension $m$ isometrically embedded in $\Rd$.
-We give to $\mc{X}$ the Riemannian structure induced by the ambient space $\Rd$, and denote the resulting volume form by $\mu$.
-
-\textcolor{red}{(TODO)} Other objects to mention, although probably some of them need only be described in the appendix.
-\begin{itemize}
-	\item Exponential map, injectivity radius, reach, Laplace-Beltrami operator.
-\end{itemize}
+Throughout this section, we will assume that the support $\mc{X}$ of the distribution $P$ is a manifold of dimension $m$ embedded in $\Rd$.
+We give to $\mc{X}$ the Riemannian structure induced by the ambient space $\Rd$, and denote the resulting volume form by $d\mu$.
+We will also assume some regularity conditions on $\mc{X}$ and $P$.
+Recall that the injectivity radius of $\mc{X}$ is the maximum value of $\delta$ such that the exponential map $\exp_x: B_m(0,\delta) \subset T_x(\mc{X}) \to B_{\mc{X}}(x,\delta) \subset \mc{X}$ is a diffeomorphism for all $x \in \mc{X}$.
+\begin{enumerate}[label=(A\arabic*)]
+	\setcounter{enumi}{4}
+	\item 
+	\label{asmp:domain_manifold} The manifold $\mc{X}$ is closed, connected, smooth and boundaryless. Additionally, the injectivity radius of $\mc{X}$ is at least $i_0 > 0$.
+	\item 
+	\label{asmp:density_manifold} The distribution $P$ admits a density $p$ with respect to the volume form $d\mu$. The density is bounded away from $0$ and $\infty$,
+	\begin{equation*}
+	0 < \frac{1}{\rho} < p(x) \leq \rho < \infty
+	\end{equation*}
+	for all $x \in \mc{X}$.
+\end{enumerate}
+The definition of the Sobolev space $H^s(P)$, and the normed ball $H^s(P;M)$ are the same, mutatis mutandi, as when $\mc{X}$ is full dimensional.
 
 \subsection{Error rates under the manifold hypothesis}
+We now show that Laplacian eigenmaps achieves the faster minimax rates under the manifold hypothesis, when the regression function $f_0 \in H^1(P;M)$ or $f_0 \in H^2(P;M)$.
+This section will proceed in a similar fashion to~\ref{sec:higher_order_sobolev_classes}, except with the ambient dimension $d$ replaced by intrinsic dimension $m$.
+
+\paragraph{Estimation.}
+To obtain a minimax optimal estimator, we choose the graph radius $\varepsilon$ and number of eigenvectors $K$ as in~\ref{asmp:parameters_estimation_ho}, except with ambient dimension $d$ replaced by the intrinsic dimension $m$.
+
+\paragraph{Testing.}
 

### 2021-03-02 notes/weekly_notes/12.24.20.NotesForWeek.tex
More progress made on the draft of Laplacian Eigenmaps paper.

@@
 We denote $T_{\varepsilon,P_n}$ by $T_{\varepsilon,n}$, and $d_{P_n}$ by $d_n$.
-For convenience, we will assume that $\eta$ is normalized, $\int_{0}^{\infty} \eta(\|z\|) \,dz = 1$, and we will denote $K(\cdot) = (\eta(\cdot))^2$.
+For convenience, we will assume that $\eta$ is normalized, $\int_{0}^{\infty} \eta(s) \,ds = 1$, and we will denote $K(\cdot) = (\eta(\cdot))^2$.
 
-The estimate we will consider is a kernel smoother passed over $\wh{\theta}$,
+The estimated function $\wh{f}$ we will consider is a kernel smoother passed over $\wh{\theta}$,
 \begin{equation}
@@
 \begin{lemma}
-	Suppose $f^{\ast} \in H_0^{s}(\Omega)$, $p \in C^{r}(\Omega)$ for $r = s - 1$, and that $\eta$ is an order-$s$ kernel. Then
+	Suppose $f^{\ast} \in H_0^{s}(\Omega)$, $p \in C^{r}(\Omega)$ for $r = s - 1$, and that $\eta$ is an order-$s$ kernel.
 	\begin{equation*}
-	\|T_{P,\varepsilon}f^{\ast} - f^{\ast}\|_P^2 \leq C \varepsilon^{2s}.
+	\|T_{\varepsilon,P}f^{\ast} - f^{\ast}\|_P^2 \leq C \varepsilon^{2s}.
 	\end{equation*}
@@
 	\label{pf:kernel_smoothing_laplacian_eigenmaps_0.5}
-	\|T_{\varepsilon,n}f^{\ast}_{1:n} - f^{\ast}\|_P^2 \leq \frac{1}{\delta}\frac{\varepsilon^2}{n\varepsilon^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{\varepsilon,P}f^{\ast} - f^{\ast}\|_P^2,
+	\|T_{\varepsilon,n}f^{\ast}_{1:n} - f^{\ast}\|_P^2 \leq \frac{C}{\delta}\frac{\varepsilon^2}{n\varepsilon^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{\varepsilon,P}f^{\ast} - f^{\ast}\|_P^2,
 	\end{equation}
@@
 
-	The left hand side of~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is equal to $\|T_{\varepsilon,n}(\wh{\theta}_{1:n} - f^{\ast}_{1:n})\|_P$, and~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is implied by showing that the operator $T_{\varepsilon,n}: L^2(X) \to L^2(\Omega)$ is (up to constant) a contraction, meaning that for any $u \in L^2(X)$,
+	The left hand side of~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is equal to $\|T_{\varepsilon,n}(\wh{\theta}_{1:n} - f^{\ast}_{1:n})\|_P$, and~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is implied by showing that the operator $T_{\varepsilon,n}: L^2(X) \to L^2(\Omega)$ is (up to a constant) a contraction, meaning that for any $u \in L^2(X)$,
 	\begin{equation}
@@
 	\begin{equation*}
-	\Bigl(\bigl(T_{n,\varepsilon}u\bigr)(x)\Bigr)^2 \leq \frac{1}{d_n(x) n} \sum_{i = 1}^{n} (u_i)^2 \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
+	\Bigl(\bigl(T_{n,\varepsilon}u\bigr)(x)\Bigr)^2 \leq \frac{1}{d_n(x) n} \sum_{i = 1}^{n} \bigl(u(X_i)\bigr)^2 \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
 	\end{equation*}
@@
 	\begin{align*}
-	T_{n,\varepsilon}(\wh{\theta} - \varepsilon)(x) & = \frac{1}{d_n(x)n} \sum_{i = 1}^{n} \bigl(\wh{\theta}_i - f^{\ast}(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr) \\
-	& = \frac{d_n(x;K^{1/2})}{d_n(x)}\frac{1}{d_n(x;K^{1/2})n} \sum_{i = 1}^{n} \bigl(\wh{\theta}_i - f^{\ast}(X_i)\bigr) \mathrm{sgn}\biggl(\eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr)\biggr) K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
+	T_{n,\varepsilon}(u)(x) & = \frac{1}{d_n(x)n} \sum_{i = 1}^{n} \bigl(u(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr) \\
+	& = \frac{d_n(x;K^{1/2})}{d_n(x)}\frac{1}{d_n(x;K^{1/2})n} \sum_{i = 1}^{n} \bigl(u(X_i)\bigr) \mathrm{sgn}\biggl(\eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr)\biggr) K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
 	\end{align*}
@@
 	\begin{align*}
-	\bigl(T_{n,\varepsilon}(\wh{\theta} - \varepsilon)(x)\bigr)^2 & \leq \biggl(\frac{d_n(x;K^{1/2})}{d_n(x)}\biggr)^2 \cdot \frac{1}{d_n(x;K^{1/2})n}\sum_{i = 1}^{n} \bigl(\wh{\theta}_i - f^{\ast}(X_i)\bigr)^2 K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr) \\
-	& \leq \frac{(1 + o_P(1))\|\eta\|_{\infty}}{\alpha} \cdot \frac{1}{d_n(x;K^{1/2})n}\sum_{i = 1}^{n} \bigl(\wh{\theta}_i - f^{\ast}(X_i)\bigr)^2 K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr),
+	\bigl(T_{n,\varepsilon}u(x)\bigr)^2 & \leq \biggl(\frac{d_n(x;K^{1/2})}{d_n(x)}\biggr)^2 \cdot \frac{1}{d_n(x;K^{1/2})n}\sum_{i = 1}^{n} \bigl(u(X_i)\bigr)^2 K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr) \\
+	& \leq \frac{(1 + o_P(1))\|\eta\|_{\infty}^2}{\alpha^2} \cdot \frac{1}{d_n(x;K^{1/2})n}\sum_{i = 1}^{n} \bigl(u(X_i)\bigr)^2 K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr),
 	\end{align*}
-	with the latter inequality following by Lemma~\ref{lem:measure_of_ball_estimate}. From here the proof proceeds along similar lines to the case where $\eta$ is non-negative.
+	with the latter inequality following by Lemma~\ref{lem:measure_of_ball_estimate} and assumption~\ref{asmp:kernel}.
+From here the proof proceeds along similar lines to the case where $\eta$ is non-negative.
 
@@
 	\begin{equation*}
-	L_{\varepsilon,n}u(x) := T_{\varepsilon,n}u(x) - u(x) = \frac{1}{d_n(x)n} \sum_{i = 1}^{n} \bigl(u(x) - u(x_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr)
+	L_{\varepsilon,n}u(x) := T_{\varepsilon,n}u(x) - u(x) = \frac{1}{d_n(x)n} \sum_{i = 1}^{n} \bigl(u(x) - u(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr)
 	\end{equation*}
@@
 	\begin{equation*}
-	\wt{L}_{\varepsilon,n}u(x) := \frac{d_n(x)}{d_P(x)}L_{\varepsilon,n}u(x) = \frac{1}{d_P(x)n} \sum_{i = 1}^{n} \bigl(u(x) - u(x_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
+	\wt{L}_{\varepsilon,n}u(x) := \frac{d_n(x)}{d_P(x)}L_{\varepsilon,n}u(x) = \frac{1}{d_P(x)n} \sum_{i = 1}^{n} \bigl(u(x) - u(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
 	\end{equation*}
-	We have already shown that $d_n(x)/d_p(x) = (1 + o_P(1))$, and so we will focus our analysis on upper bounding $\Ebb[\|\wt{L}_{\varepsilon,n}f^{\ast}\|_P^2]$.
+	We have already shown that $d_n(x)/d_P(x) = (1 + o_P(1))$, and so we will focus our analysis on upper bounding $\Ebb[\|\wt{L}_{\varepsilon,n}f^{\ast}\|_P^2]$.
 We begin by providing expressions for the pointwise mean and variance of $\wt{L}_{\varepsilon,n}f^{\ast}(x)$:
@@
 	\end{align*}
-	where $E_{\varepsilon}(f^{\ast}) := \int_{\Omega} \int_{\Omega} (f^{\ast}(x') - f^{\ast}(x))^2 K(\|x' - x\|/\varepsilon) p(x') p(x) ,\dx' \,dx$, and we have shown in our AISTATS paper that $E_{\varepsilon}(f^{\ast}) \leq C \varepsilon^{d + 2} \|f\|_{H^1(\Omega)}^2$, justifying the second inequality.
+	where $E_{\varepsilon}(f^{\ast}) := \int_{\Omega} \int_{\Omega} (f^{\ast}(x') - f^{\ast}(x))^2 K(\|x' - x\|/\varepsilon) p(x') p(x) ,\dx' \,dx$, and we have shown in our AISTATS paper that $E_{\varepsilon}(f^{\ast}) \leq C \varepsilon^{d + 2} |f|_{H^1(\Omega)}^2$, justifying the second inequality.
+Consequently,
+	\begin{align*}
+	\|T_{n,\varepsilon}f^{\ast}_{1:n} - f^{\ast}\|_P^2 \leq (1 + o_P(1))^2 \|\wt{L}_{\varepsilon,n}f^{\ast}\|_P^2 & \leq \frac{\bigl(1 + o_P(1)\bigr)^2}{\delta} \mathbb{E}\bigl[\|\wt{L}_{\varepsilon,n}f^{\ast}\|_P^2\bigr] \\
+	& \leq \frac{\bigl(1 + o_P(1)\bigr)^2}{\delta}\biggl(\|T_{P,\varepsilon}f^{\star} - f^{\star}\|_{P}^2 + \frac{C\varepsilon^2}{p_{\min}^2 \alpha^2 \varepsilon^{d}} |f|_{H^1(\Omega)}^2\biggr)
+	\end{align*}
+	with probability at least $1 - \delta$, establishing~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0.5}.
 \end{proof}

### 2021-03-03 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Added estimate on bias term of kernel smoothing when d = 1.

@@
 \subsection{Spectral series methods for Sobolev Spaces}
-Spectral series estimators are particularly designed for estimation in Sobolev spaces.
+Spectral series estimators---and tests using their $\Leb^2$-norm---are particularly designed for regression in Sobolev spaces.
 Let $f_1,f_2,\ldots$ be the unit norm eigenfunctions of $\Delta_P$.
@@
 We show that $\wh{f}$ and a test based on $\wh{T}$ are minimax optimal, for all values of $d$, and under no additional assumptions on the data generating process, i.e.
-on $P$ or $f_0$.
+on either $P$ or $f_0$.
 
@@
 	\item 
-	The graph radius $\varepsilon$ and number of eigenvectors $K$ satisfies
+	\label{asmp:parameters_estimation_ho}
+	The graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy
 	\begin{equation*}
@@
 For instance, it is not a Reproducing Kernel Hilbert Space, nor does it embed into any H\"{o}lder space.
-As a result, certain versions of the nonparametric regression problem are ill-posed---for instance when loss is measured in $\Leb^{\infty}$ norm, or when the design points $X$ are assumed to be fixed.
+As a result, certain versions of the nonparametric regression problem are ill-posed---for instance when loss is measured in $\Leb^{\infty}$ norm, or when the design points $\{X_1,\ldots,X_n\}$ are assumed to be fixed.
 Likewise, certain estimators are ``off the table'', most notably RKHS-based methods such as smoothing splines.
-However, at these scales spectral series estimators still obtain the optimal rates for random design regression measured in $\Leb^2(P)$ norm.
+Nevertheless, for random design regression with error measured in $\Leb^2(P)$-norm, spectral series estimators obtain the minimax rates for all values of $s$ and $d$.
 Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} shows that the same is true with respect to Laplacian eigenmaps, with error measured in $\Leb^2(P_n)$ norm.
@@
 When $4 < d \leq 4s$, we get the rate~\textcolor{red}{(...)} Recall that the rate $n^{-4s/(4s + d)}$ is optimal whenever $d \leq 4s$.
-Therefore there are scalings of $s$ and $d$ for which we do not get the optimal rate.
+Thus when $4s > d$, our upper bounds do not imply that $\varphi$ is an optimal test.
 We do not know whether this is due to looseness in our proof techniques, or a problem with the test itself, and leave this for future work.
@@
 where $L_{P,\varepsilon}$ is a non-local approximation to $\Delta_P$, 
-\begin{equation*}
-L_{P,\varepsilon}f(x) := \int_{\mc{X}}\bigl(f(z) - f(x)\bigr) \eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dP(x).
-\end{equation*}
+\begin{equation}
+\label{eqn:nonlocal_laplacian}
+L_{P,\varepsilon}f(x) := \frac{1}{\varepsilon^{d + 2}}\int_{\mc{X}}\bigl(f(z) - f(x)\bigr) \eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dP(x).
+\end{equation}
 In the second step, we show that $E_{\varepsilon}^{(s)}(f)$ is in turn upper bounded by a constant times $D_s(f) := \|f\|_{H^s(P)}^2$.
@@
 In words, we show that a $j$th-order iterated difference operator approximates a $2j$-th (or $2j + 1$-th) order differential operator.
-It is in order to establish this fact that we require on the one hand $p \in C^{r}(\mc{X};M)$, and on the other hand that $f$ be zero-trace.
+In to establish this fact, we require that on the one hand $p \in C^{r}(\mc{X};M)$, and on the other hand $f$ be zero-trace.
 
@@
 
+\begin{enumerate}[label=(A\arabic*)]
+	\setcounter{enumi}{5}
+	\item 
+	\label{asmp:parameters_estimation_manifold}
+	The graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy
+	\begin{equation}
+	\label{eqn:radius_estimation_manifold}
+	C_0 \cdot \max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/m}, n^{-1/(2r + m)}\biggr\} \leq \varepsilon \leq \min\{i_0, K^{-1/m}\}
+	\end{equation}
+	and
+	\begin{equation*}
+	K \asymp (M^2n)^{2m/(2s + m)}
+	\end{equation*}
+\end{enumerate}
+
+\begin{theorem}
+	\label{thm:laplacian_eigenmaps_estimation_manifold}
+	Suppose $f_0 \in H^s(\mc{X},M)$, for $s \leq 2$. Under appropriate conditions, with probability at least $1 - \delta - \textcolor{red}{(?)}$,
+	\begin{equation}
+	\label{eqn:laplacian_eigenmaps_estimation_manifold}
+	\|\wh{f} - f_0\|_n^2 \lesssim M^{2m/(2s + m)}n^{-2s/(2s + m)}
+	\end{equation}
+\end{theorem}
+
 \paragraph{Testing.}
+To construct a minimax optimal test using $\wh{T}$, we choose $\varepsilon$ and $K$ as in~\ref{asmp:parameters_testing_fo}, except with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
+\begin{enumerate}[label=(A\arabic*)]
+	\setcounter{enumi}{5}
+	\item 
+	\label{asmp:parameters_testing_manifold}
+	The graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy~\eqref{eqn:radius_estimation_manifold}, for $r = 0$. Additionally, there exist constants $c_2$ and $C_2$ such that
+	\begin{equation*}
+	c_2 (M^2n)^{2m/(4s + m)} \leq K \leq C_2 (M^2n)^{2m/(4s + m)}
+	\end{equation*}
+\end{enumerate}
+
+\begin{theorem}
+	\label{thm:laplacian_eigenmaps_testing_manifold}
+	Suppose $f_0 \in H^1(\mc{X};M)$.
+If the Laplacian eigenmaps test $\varphi$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_testing_manifold}, then the following statement holds for any $\delta \in (0,1)$: if $f_0$ satisfies
+	\begin{equation}
+	\label{eqn:laplacian_eigenmaps_testing_criticalradius_manifold}
+	\|f_0\|_P^2 \geq \frac{C}{\delta} \max\biggl\{\frac{1}{\delta}M^{2m/(4 + m)}n^{-4/(4 + m)}, M n^{-p(m)}\biggr\},
+	\end{equation}
+	then $\varphi$ has Type II error upper bounded by 
+	\begin{equation}
+	\Pbb_{f_0}(\varphi = 0) \leq \delta + \textcolor{red}{(...)}
+	\end{equation}
+\end{theorem}
+
+\begin{itemize}
+	\item \textbf{Proof strategy.} The proofs of Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold} proceed in a similar manner to the full-dimensional setting.
+	\item \textbf{Restriction on number of derivatives.} Unlike in the full-dimensional case, Theorem~\ref{thm:laplacian_eigenmaps_estimation_manifold} holds only when $s \leq 2$.
+Roughly speaking, we need this restriction in order to show that $L_{P,\varepsilon}^sf(x)$ is a good approximation to $\Delta_P^sf(x)$.
+Otherwise, the error incurred by using Euclidean distance rather than geodesic distance in~\eqref{eqn:nonlocal_laplacian} will dominate.
+\end{itemize}
 
@@
 
-Of course, the Laplacian eigenmaps estimator is only defined at the observed design points $X_1,\ldots,X_n$, and to measure its error in $L^2(P)$ norm we must first give a means of extending it to the rest of $\Xset$.
+Of course, the Laplacian eigenmaps estimator is only defined at the observed design points $X_1,\ldots,X_n$, and to measure its error in $L^2(P)$ norm we must first extend it to the rest of $\Xset$.
 We propose a simple method, essentially kernel smoothing, to do the job, which can be applied to any estimator defined at the design points, including Laplacian eigenmaps.

### 2021-03-03 notes/weekly_notes/12.24.20.NotesForWeek.tex
Added estimate on bias term of kernel smoothing when d = 1.

@@
 Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data (on the order of $1/(n\varepsilon^d)$) by a factor of $\varepsilon^2$.
-The bias term, on the other hand, is standard, and in particular the following estimate of the bias can be obtained.
+The bias term, on the other hand, is standard (at least when $f^{\ast} \in C_0^s(\Omega)$).
+In particular the following estimate of the bias can be obtained.
 \begin{lemma}
-	Suppose $f^{\ast} \in H_0^{s}(\Omega)$, $p \in C^{r}(\Omega)$ for $r = s - 1$, and that $\eta$ is an order-$s$ kernel.
+	\label{lem:kernel_smoothing_bias}
+	Suppose $f^{\ast} \in H_0^{s}(\Omega)$, $p \in C^{r}(\Omega)$ for $r = s - 1$, and that $\eta$ is an order-$s$ kernel. It follows that
 	\begin{equation*}
@@
 \end{lemma}
-\begin{proof}
-	\textcolor{red}{(TODO)}
+\begin{proof}[Proof (when dimension $d = 1$).]
 \end{proof}
@@
 
-This one I frankly have no thoughts on how to analyze. I don't even understand it all that well, as evidence by the fact that I can't name it.
+This one I frankly have no thoughts on how to analyze. I don't even understand it all that well, as evidenced by the fact that I can't name it.
 
@@
 
+\section{Miscellaneous}
+I put things here which did not belong in any other section.
+
+\subsection{Proof of Lemma~\ref{lem:kernel_smoothing_bias}, $d = 1$ and $p$ uniform}
+Without loss of generality take the domain $\Omega = (0,1)$. Recall that we are interested in upper bounding
+\begin{equation*}
+\|T_{P,\varepsilon}u - u\|_{P}^2 = \int_{\Omega} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x) \leq C\varepsilon^{2k}
+\end{equation*}
+where $L_{P,\varepsilon}$ is a non-local approximation to the continuum weighted Laplacian, given by
+\begin{equation*}
+L_{P,\varepsilon}u(x) := \frac{1}{d_P(x)}\int_{\Omega} \bigl(u(z) - u(x)\bigr) \eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dP(z).
+\end{equation*}
+We begin with a pointwise analysis of $L_{p,\varepsilon}u(x)$ at a given $x \in \Omega$.
+We may assume without loss of generality that $u \in C_0^{\infty}(\Omega)$, since $C_0^{\infty}(\Omega)$ is dense in $H_0^k(\Omega)$ and $\|T_{P,\varepsilon}u - u\|_P^2$ is continuous with respect to $\|\cdot\|_{H_0^k}$.
+Since $u$ is smooth, for every $x,z \in \Omega$, we have the $k$th-order Taylor expansion
+\begin{equation*}
+u(z) - u(x) = \sum_{j = 1}^{k - 1} (z - x)^{j} u^{(j)}(x)  + \frac{(z - x)^k}{(k - 1)!}\int_{0}^{1}(1 - t)^{k - 1}u^{(k)}(x + t(z - x))\,dt.
+\end{equation*}
+As a result,
+\begin{equation}
+\label{pf:kernel_smoothing_bias_1d_1}
+\begin{aligned}
+L_{P,\varepsilon}u(x) = \frac{1}{d_P(x)} & \Biggl\{\sum_{j = 1}^{k - 1} u^{(j)}(x) \int_{\Omega}(z - x)^{j}\eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr)\,dz + \\ 
+& \frac{1}{(k - 1)!} \int_{\Omega} \int_{0}^{1} (1 - t)^k (z - x)^k u^{(k)}(x + t(z - x)) \eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dt \,dz \Biggr\}
+\end{aligned}
+\end{equation}
+The way in which we use~\eqref{pf:kernel_smoothing_bias_1d_1} will depend on whether $x \in \Omega$ satisfies $B(x,\varepsilon) \subset \Omega$ or not.
+Let $\Omega_{\varepsilon} := \{x: B(x,\varepsilon) \in \Omega\}$ and let $\partial_{\varepsilon}\Omega := \Omega \setminus \Omega_{\varepsilon}$.
+Clearly
+\begin{equation*}
+\|L_{P,\varepsilon}u\|_{P}^2 = \underbrace{\int_{\Omega_{\varepsilon}} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_1} + \underbrace{\int_{\partial_{\varepsilon}\Omega} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_2}
+\end{equation*}
+and we now separately upper bound $I_1$ and $I_2$.
+
+\paragraph{Upper bound on $I_1$.}
+If $x \in \Omega_{\varepsilon}$ then all $z \in B(x,\varepsilon)$ also belong to $\Omega$. For one thing, this implies
+\begin{equation*}
+d_P(x) = \int_{\Omega} \eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dz = \varepsilon \cdot \int_{B(0,1)}  \eta(v) \,dv = \epsilon.
+\end{equation*}
+Addtionally, we have that all $j = 1,\ldots,k - 1$, 
+\begin{equation*}
+\int_{\Omega}(z - x)^{k}\eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dz = \varepsilon^j \cdot \int_{B(0,1)} v^j \eta(v) \,dv = 0,
+\end{equation*}
+where the second equality follows because $\eta$ is a $k$th-order kernel.
+Consequently $L_{P,\varepsilon}$ depends only on the $k$th derivative of $u$: changing variables to $v = (z - x)/\varepsilon$, we have
+\begin{align*}
+L_{P,\varepsilon}u(x) & = \frac{1}{\varepsilon} \cdot \frac{1}{(k - 1)!} \int_{0}^{1} \int_{\Omega} (1 - t)^k (z - x)^k u^{(k)}(x + t(z - x)) \eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dz \,dv \\
+& = \frac{\varepsilon^{k}}{(k - 1)!} \int_{0}^{1} \int_{B(0,1)} (1 - t)^k v^k u^{(k)}(x + t\varepsilon v) \eta(v) \,dv \,dt.
+\end{align*} 
+This identity along with Jensen's inequality implies that
+\begin{equation*}
+\Bigl(L_{P,\varepsilon}u(x)\Bigr)^2 \leq \biggl(\frac{\varepsilon^{k}}{(k - 1)!}\biggr)^2 \cdot  \int_{0}^{1} \int_{B(0,1)} v^{2k} \bigl(u^{(k)}(x + t\varepsilon v)\bigr)^2 \eta(|v|) \,dv \,dt.
+\end{equation*}
+Integrating over all $x \in \Omega_{\varepsilon}$ gives an upper bound on $I_1$. 
+From here, by changing the order of integration, substituting $\wt{x} = x + t\varepsilon v$, and noting that $\wt{x} \in \Omega$, we arrive at the desired bound:
+\begin{align*}
+I_1 & \leq \biggl(\frac{\varepsilon^{k}}{(k - 1)!}\biggr)^2 \int_{0}^{1} \int_{B(0,1)} \int_{\Omega_{\varepsilon}} v^{2k} \bigl(u^{(k)}(x + t\varepsilon v)\bigr)^2 \eta(|v|) \,dx \,dv \,dt \\
+& \leq \biggl(\frac{\varepsilon^{k}}{(k - 1)!}\biggr)^2 \int_{0}^{1} \int_{B(0,1)} \int_{\Omega} v^{2k} \bigl(u^{(k)}(\wt{x})\bigr)^2 \eta(|v|) \,d\wt{x} \,dv \,dt \\
+& = \biggl(\frac{\varepsilon^{k}}{(k - 1)!}\biggr)^2 |u|_{H^k(\Omega)}^2 \sigma_{\eta}.
+\end{align*}
+
+\paragraph{Upper bound on $I_2$.}
+
 \end{document}

### 2021-03-03 manuscripts/laplacian_smoothing/arXiv/appendix.tex
Updated arXiv version of draft to be in-sync with AISTATS camera-ready submission.

@@
-\noindent In this appendix, we provide proofs of all results found in the main text.
+\noindent In the appendix, we provide proofs of all results found in the paper ``Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs''.
 Our main theorems (Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing_manifold}) all follow the same general proof strategy, centered on conditioning.
-In Section~\ref{sec:fixed_graph_error_bounds}, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to (functionals of) the graph $G$, and allow us to upper bound the error of $\wh{f}$ and $\wh{\varphi}$ conditional on the design $\mathbf{X} = x$.
+In Section~\ref{sec:fixed_graph_error_bounds}, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to (functionals of) the graph $G$, and allow us to upper bound the error of $\wh{f}$ and $\wh{\varphi}$ conditional on the design $\{X_1,\ldots,X_n\} = \{x_1,\ldots,x_n\}$.
 In Sections~\ref{sec:graph_sobolev_seminorm}, \ref{sec:graph_eigenvalues}, \ref{sec:empirical_norm}, and \ref{sec:manifold} we give all our necessary probabilistic estimates on these functionals, for the particular random neighborhood graph $G = G_{n,r}$.
@@
 
+\paragraph{Pointwise evaluation of Sobolev functions.}
+First, however, as promised in our main text we clarify what is meant by pointwise evaluation of the regression function $f_0$.
+Strictly speaking, each $f \in H^1(X)$ is really an equivalence class, defined only up to sets of Lebesgue measure 0.
+In order to make sense of the evaluation $x \mapsto f(x)$, one must therefore pick a representative $f^{\star} \in f$.
+When $d = 1$, this is resolved in a standard way---since $H^1(\mc{X})$ embeds continuously into $C^0(\mc{X})$, there exists a continuous version of every $f \in H^1(\mc{X})$, and we take this continuous version as the representative $f^{\star}$.
+On the other hand, when $d \geq 2$, the Sobolev space $H^1(\mc{X})$ does not continuously embed into $C^0(\mc{X})$, and we must choose representatives in a different manner.
+In this case we let $f^{\star}$ be the precise representative \citep{evans15}, defined pointwise at points $x \in \mc{X}$ as
+\begin{equation*}
+f^{\star}(x) = 
+\begin{dcases*}
+\lim_{\varepsilon \to 0} \frac{1}{\nu(B(x,\varepsilon))} \int_{B(x,\varepsilon)} f(x) dx,&~~\textrm{if the limit exists,} \\
+0,&~~\textrm{otherwise.}
+\end{dcases*}
+\end{equation*}
+Note that when $d = 1$, the precise representative of any $f \in H^1(\mc{X})$ is continuous.
+
+Now we explain why the particular choice of representative is not crucial, using the notion of a Lebesgue point.
+Recall that for a locally Lebesgue integrable function $f$, a given point $x \in \mc{X}$ is a \emph{Lebesgue point} of $f$ if the limit of $1/(\nu(B(x,\varepsilon)))\int_{B(x,\varepsilon)} f(x) dx$ as $\varepsilon \to 0$ exists, and satisfies
+\begin{equation*}
+\lim_{\varepsilon \to 0} \frac{1}{\nu\bigl(B(x,\varepsilon)\bigr)} \int_{B(x,\varepsilon)} f(x) dx = f(x).
+\end{equation*}
+Let $E$ denote the set of Lebesgue points of $f$.
+By the Lebesgue differentiation theorem \citep{evans15}, if $f \in L^1(\mc{X})$ then almost every $x \in \mc{X}$ is a Lebesgue point, $\nu(\mc{X}\setminus E) = 0$.
+Since $f_0 \in H^1(\mc{X}) \subseteq \Leb^1(\mc{X})$, we can conclude that any function $g_0 \in f_0$ disagrees with the precise representative $f_0^{\star}$ only on a set of Lebesgue measure 0.
+Moreover, since we always assume the design distribution $P$ has a continuous density, with probability $1$ it holds that $g_0(X_i) = f_0^{\star}(X_i)$ for all $i = 1,\ldots,n$.
+This justifies the notation $f_0(X_i)$ used in the main text.
+
+\vfill
+
+\clearpage
+
 \section{Graph-dependent error bounds}
 \label{sec:fixed_graph_error_bounds}
+
 In this section, we adopt the fixed design perspective; or equivalently, condition on $X_i = x_i$ for $i = 1,\ldots,n$.
@@
 
-To prove Lemma~\ref{lem:graph_sobolev_seminorm}, we will show that for any $f \in H^1(\Xset)$,
-\begin{equation*}
-\Ebb\Bigl[f^{\top} \Lap_{n,r} f\Bigr] \leq C_2 n^2 r^{d + 2} \|f\|_{H^1(\Xset)}^2,
-\end{equation*}
-whence the claim follows immediately by Markov's inequality (recall that $\Lap_{n,r}$ is positive semi-definite, and therefore $f^{\top} \Lap_{n,r} f$ is a non-negative random variable).
+Throughout this proof, we will assume that $f \in H^1(\Xset)$ has zero-mean, meaning $\int_{\Xset} f(x) \,dx = 0$.
+This is without loss of generality---assuming for the moment that \eqref{eqn:graph_sobolev_seminorm} holds for zero-mean functions, for any $f \in H^1(\Xset)$, taking $a = \int_{\Xset} f(x) \,dx$ and $g = f - a$, we have that
+\begin{equation*}
+f^{\top} L_{n,r} f = g^{\top} L_{n,r} g \leq \frac{C_2}{\delta} n^2 r^{d + 2} |g|_{H^1(\mc{X})}^2 = \frac{C_2}{\delta} n^2 r^{d + 2} |f|_{H^1(\mc{X})}^2.
+\end{equation*} 
+Now, for any zero-mean function $f \in H^1(\mc{X})$ it follows by the Poincare inequality (see Section 5.8, Theorem 1 of \citet{evans10}) that $\|f\|_{H^1(\mc{X})}^2 \leq C_8 |f|_{H^1(\mc{X})}^2$, for some constant $C_8$ that does not depend on $f$.
+Therefore, to prove Lemma~\ref{lem:graph_sobolev_seminorm}, it suffices to show that
+\begin{equation*}
+\Ebb\Bigl[f^{\top} \Lap_{n,r} f\Bigr] \leq C n^2 r^{d + 2} \|f\|_{H^1(\Xset)}^2,
+\end{equation*}
+since the high-probability upper bound then follows immediately by Markov's inequality.
+(Recall that $\Lap_{n,r}$ is positive semi-definite, and therefore $f^{\top} \Lap_{n,r} f$ is a non-negative random variable).
 
@@
 
-Take $\Omega$ to be an arbitrary bounded open set such that $B(x,r) \subset \Omega$ for all $x \in \mc{X}$.
+Now, take $\Omega$ to be an arbitrary bounded open set such that $B(x,c_0) \subseteq \Omega$ for all $x \in \mc{X}$.
 For the remainder of this proof, we will assume that (i) $f \in H^1(\Omega)$ and additionally (ii) $\|f\|_{H^1(\Omega)} \leq C_5 \|f\|_{H^1(\mc{X})}$ for a constant $C_5$ that does not depend on $f$.
-This is without loss of generality, since by Theorem~1 of~\citet{evans10} there exists an extension operator $E: H^1(\mc{X}) \to H^1(\Omega)$ for which the extension $Ef$ satisfies both (i) and (ii).
+This is without loss of generality, since by Theorem~1 in Chapter 5.4 of~\citet{evans10} there exists an extension operator $E: H^1(\mc{X}) \to H^1(\Omega)$ for which the extension $Ef$ satisfies both (i) and (ii).
 Additionally, we will assume $f \in C^{\infty}(\Omega)$.
-Again, this is without loss of generality, as $C^{\infty}(\Omega)$ is dense in $H^1(\Omega)$ and the expectation on the right hand side of~\eqref{pf:first_order_graph_sobolev_seminorm_1} is continuous in $H^1(\Omega)$.
+Again, this is without loss of generality, as $C^{\infty}(\Omega)$ is dense in $H^1(\Omega)$ and the expectation on the right hand side of \eqref{pf:first_order_graph_sobolev_seminorm_1} is continuous in $H^1(\Omega)$.
 The reason for dealing with a smooth extension $f \in C^{\infty}(\Omega)$ is so that we can make sense of the following equality for any $x$ and $x'$ in $\mc{X}$:
@@
 \end{equation*}
-proving the claim of Lemma~\ref{lem:graph_sobolev_seminorm} upon taking $C_2 := C_5 \sigma_K p_{\max}^2$ in the statement of the lemma.
+proving the claim of Lemma~\ref{lem:graph_sobolev_seminorm} upon taking $C_2 := C_8C_5 \sigma_K p_{\max}^2$ in the statement of the lemma.
 
 \subsection{Stronger bounds under Lipschitz assumption}
+
 Suppose $f$ satisfies $|f(x') - f(x)| \leq M \|x - x\|$ for all $x,x' \in \mc{X}$.
@@
 \end{proposition}
-See Lemma 28 of~\citet{dunlop2020} for a proof that~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density} imply Weyl's Law.
+See Lemma 28 of~\citet{dunlop2020} for a proof that~\ref{asmp:domain} and~\ref{asmp:density} imply Weyl's Law.
+
 \paragraph{Proof of Lemma~\ref{lem:neighborhood_eigenvalue}.}
@@
 \paragraph{Proof of Lemma~\ref{lem:empirical_norm_sobolev}.}
-To establish~\eqref{eqn:empirical_norm_sobolev}, it is sufficient to show that
+In this proof, we will find it more convenient to deal with the parameterization $b = 1/\delta$. To establish \eqref{eqn:empirical_norm_sobolev}, it is sufficient to show that
 \begin{equation*}
@@
 \paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}.}
-We have shown that the inequalities~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} are satisfied with probability at least $1 - \delta - C_1n\exp(-c_1nr^d)$.
-From~\eqref{eqn:neighborhood_eigenvalue} and Lemma~\ref{lem:variance_term_estimation}, we have that
-\begin{equation*}
-\sum_{k = 1}^{n}\biggl(\frac{1}{\rho \lambda_k + 1}\biggr)^{2} \geq 1 + \sum_{k = 2}^{n}\biggl(\frac{1}{C_3 \wt{\rho} k^{2/d} + 1}\biggr)^{2} \geq 1 + \frac{1}{4} \min\Bigl\{\floor{\frac{\wt{\rho}^{-d/2}}{C_3^{d/2}}} - 1 \vee 0,n\Bigr\} \geq \min\Bigl\{\frac{\wt{\rho}^{-2/d}}{8C_3^{2/d}}, \frac{n}{4} \Bigr\}
-\end{equation*}
-Therefore by Lemma~\ref{lem:ls_fixed_graph_estimation} along with~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue}, with probability at least $1 - \delta - C_1n\exp(-c_1nr^d) - \exp(-\min\{\wt{\rho}^{-d/2}/C_3^{d/2},n\}/8)$ it follows that,
+We have shown that the inequalities \eqref{eqn:graph_sobolev_seminorm} and \eqref{eqn:neighborhood_eigenvalue} are satisfied with probability at least $1 - \delta - C_1n\exp(-c_1nr^d)$, and throughout this proof we take as granted that both of these inequalities hold.
+
+Now, set $\wt{\rho} = M^{-4/(2 + d)}n^{-2/(2 + d)}$ as prescribed in Theorem~\ref{thm:laplacian_smoothing_estimation1}, and note that $\wt{\rho}^{-d/2} \leq n$ is implied by the assumption $M \leq n^{1/d}$.
+Therefore  from \eqref{eqn:neighborhood_eigenvalue} and Lemma~\ref{lem:variance_term_estimation}, it follows that
+\begin{equation*}
+\sum_{k = 1}^{n}\biggl(\frac{1}{\rho \lambda_k + 1}\biggr)^{2} \geq 1 + \frac{1}{C_3^2}\sum_{k = 2}^{n}\biggl(\frac{1}{\wt{\rho} k^{2/d} + 1}\biggr)^{2} \geq \frac{1}{8C_3^2} \wt{\rho}^{-d/2}.
+\end{equation*}
+As a result, by Lemma~\ref{lem:ls_fixed_graph_estimation} along with \eqref{eqn:graph_sobolev_seminorm} and \eqref{eqn:neighborhood_eigenvalue}, with probability at least $1 - \delta - C_1n\exp(-c_1nr^d) - \exp(-\wt{\rho}^{-d/2}/8C_3^2)$ it holds that,
 \begin{align}
 \|\wh{f} - f_0\|_n^2 & \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{n}\sum_{k = 2}^{n} \Biggl(\frac{1}{c_3 \wt{\rho}\min\{k^{2/d},r^{-2}\} + 1}\Biggr)^2 \nonumber \\
-& \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{n}\sum_{k = 1}^{n} \biggl(\frac{1}{c_3 \wt{\rho}k^{2/d} + 1}\biggr)^2 + \frac{10r^4}{c_3^2 \wt{\rho}^2}.
+& \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{nc_3^2}\sum_{k = 2}^{n} \biggl(\frac{1}{\wt{\rho}k^{2/d} + 1}\biggr)^2 + \frac{10r^4}{c_3^2 \wt{\rho}^2}.
 \label{pf:laplacian_smoothing_estimation1_1}
 \end{align}
-Setting $\wt{\rho} = M^{-4/(2 + d)}n^{-2/(2 + d)}$ as prescribed in Theorem~\ref{thm:laplacian_smoothing_estimation1},  we have from Lemma~\ref{lem:variance_term_estimation} that
-\begin{align*}
-\|\wh{f} - f_0\|_n^2 & \leq\frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{20\wt{\rho}^{-d/2}}{c_3^{d/2}n}  + \frac{10r^4}{c_3^2 \wt{\rho}^2} \\
-& = \biggl(\frac{C_2}{\delta} + \frac{20}{c_3^{d/2}}\biggr)M^{2d/(2+d)} n^{-2/(2 + d)} + \frac{10}{n} + \frac{10}{c_3^2}r^4M^{8/(2 + d)}n^{4/(2 + d)} \\
-& \leq \biggl(\frac{C_2}{\delta} + \frac{20}{c_3^{d/2}} + 10 + \frac{10}{c_3^2}\biggr)M^{2d/(2+d)} n^{-2/(2 + d)}
-\end{align*}
-with the last inequality following from~\ref{asmp:ls_kernel_radius_estimation} and the assumption $M \geq n^{-1/2}$.
-This completes the proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}.
-When $M < n^{-1/2}$, one can set $\wt{\rho} = 1$ and trace through the previous calculations to determine that $\|\wh{f} - f_0\|_n^2 \leq C/n$ for an appropriate choice of $C$.
-
-\paragraph{Nearly-optimal rates when $d = 4$.}
-Continuing on from~\eqref{pf:laplacian_smoothing_estimation1_1}, from Lemma~\ref{lem:variance_term_estimation} we have that
+The first term on the right hand side of \eqref{pf:laplacian_smoothing_estimation1_1} is a bias term, while the second, third, and fourth terms each contribute to the variance.
+Of these, under our assumptions the third term dominates, as we show momentarily.
+First, we use Lemma~\ref{lem:variance_term_estimation} to get an upper bound on this variance term,
+\begin{equation*}
+\sum_{k = 2}^{n}\biggl(\frac{1}{\wt{\rho}k^{2/d} + 1}\biggr)^2 \leq 4\wt{\rho}^{-d/2}.
+\end{equation*}
+Then plugging this upper bound back into \eqref{pf:laplacian_smoothing_estimation1_1}, we have that
+\begin{align*}
+\|\wh{f} - f_0\|_n^2 & \leq\frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{40\wt{\rho}^{-d/2}}{c_3^{2}n}  + \frac{10r^4}{c_3^2 \wt{\rho}^2} \\
+& = \biggl(\frac{C_2}{\delta} + \frac{40}{c_3^{2}}\biggr)M^{2d/(2+d)} n^{-2/(2 + d)} + \frac{10}{n} + \frac{10}{c_3^2}r^4M^{8/(2 + d)}n^{4/(2 + d)} \\
+& \leq \biggl(\frac{C_2}{\delta} + \frac{50}{c_3^{2}}\biggr)M^{2d/(2+d)} n^{-2/(2 + d)}
+\end{align*}
+with the last inequality following from~\ref{asmp:ls_kernel_radius_estimation} and the assumption $M \geq n^{-1/2}$. This completes the proof of Theorem~\ref{thm:laplacian_smoothing_estimation1}.
+
+\paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_estimation2}.}
+We first establish that $\wh{f}$ achieves nearly-optimal rates when $d = 4$, and then prove the sub-optimal rates when $d > 4$.
+
+\textit{Nearly-optimal rates when $d = 4$.}
+
+Continuing on from \eqref{pf:laplacian_smoothing_estimation1_1}, from Lemma~\ref{lem:variance_term_estimation} we have that
 \begin{equation*}
@@
 
-\paragraph{Suboptimal rates when $d > 4$.}
-Once again continuing on from~\eqref{pf:laplacian_smoothing_estimation1_1}, from Lemma~\ref{lem:variance_term_estimation} we have that,
-\begin{equation*}
-\|\wh{f} - f_0\|_n^2 \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{nc_3^{d/2}\wt{\rho}^{d/2}} +  \frac{10}{n^{4/d}\wt{\rho}^2c_3^2} + \frac{10r^4}{\wt{\rho}^2c_3^2}
+\textit{Suboptimal rates when $d > 4$.}
+
+Once again continuing on from \eqref{pf:laplacian_smoothing_estimation1_1}, from Lemma~\ref{lem:variance_term_estimation} we have that,
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{nc_3^{2}\wt{\rho}^{d/2}} +  \frac{10}{n^{4/d}\wt{\rho}^2c_3^2} + \frac{10r^4}{\wt{\rho}^2c_3^2}
 \end{equation*}
@@
 \paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_testing}.}
-Recall that we have shown that the inequalities~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} are satisfied with probability at least $1 - \delta - C_1n\exp(-c_1nr^d)$.
-
-Now, we would like to invoke Lemma~\ref{lem:ls_fixed_graph_testing}, and in order to do so, we must show that the inequality~\eqref{eqn:ls_fixed_graph_testing_critical_radius} is satisfied with respect to $G = G_{n,r}$.
+Let $\delta = 1/b$.
+Recall that we have shown that the inequalities \eqref{eqn:graph_sobolev_seminorm} and \eqref{eqn:neighborhood_eigenvalue} are satisfied with probability at least $1 - 1/b - C_1n\exp(-c_1nr^d)$, and throughout this proof we take as granted that both of these inequalities hold.
+
+Now, we would like to invoke Lemma~\ref{lem:ls_fixed_graph_testing}, and in order to do so, we must show that the inequality \eqref{eqn:ls_fixed_graph_testing_critical_radius} is satisfied with respect to $G = G_{n,r}$.
 First, we upper bound the right hand side of this inequality.
-By~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue},
-\begin{align*}
-\frac{2 \rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{4b}{n} \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{1/2} & \leq C_2b\wt{\rho}M^2 + \frac{4b}{n}\Biggl[1 + \biggl(\sum_{k = 2}^{n} \frac{1}{(c_3\wt{\rho}k^{2/d} + 1)^4} \biggr)^{1/2} + \frac{r^4n^{1/2}}{c_3^2\wt{\rho}^2}\Biggr] \\
-& \leq C_2b\wt{\rho}M^2 + \frac{4b}{n}\biggl(1 + \frac{\sqrt{2}}{c_3^{d/4}}\wt{\rho}^{-d/4} + \frac{r^4n^{1/2}}{c_3^2\wt{\rho}^2}\biggr) \\ 
-& \leq \Bigl(C_2 + 4 + \frac{4\sqrt{2}}{c_3^{d/4}} + \frac{4}{c_3^2}\Bigr)bM^{2d/(4 + d)} n^{-4/(4 + d)}
-\end{align*}
-where the second inequality is justfied by Lemma~\ref{lem:variance_term_testing}, and in the third inequality we set $\wt{\rho} = M^{-8/(4 + d)}n^{-4/(4 + d)}$ as prescribed by Theorem~\ref{thm:laplacian_smoothing_testing}, and use both the upper bound on $r$ assumed in~\ref{asmp:ls_kernel_radius_testing} as well as the fact that $M \geq n^{-1/2}$.
-
-Next we lower bound the left hand side of the inequality~\eqref{eqn:ls_fixed_graph_testing_critical_radius}---i.e.
+Set $\wt{\rho} = M^{-8/(4 + d)}n^{-4/(4 + d)}$ as prescribed by Theorem~\ref{thm:laplacian_smoothing_testing}, it follows from \eqref{eqn:graph_sobolev_seminorm} and \eqref{eqn:neighborhood_eigenvalue} that
+\begin{align*}
+\frac{2 \rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{4b}{n} \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{1/2} & \leq C_2b\wt{\rho}M^2 + \frac{4b}{n}\Biggl[1 + \frac{1}{c_3^2}\biggl(\sum_{k = 2}^{n} \frac{1}{(\wt{\rho}k^{2/d} + 1)^4} \biggr)^{1/2} + \frac{r^4n^{1/2}}{c_3^2\wt{\rho}^2}\Biggr] \\
+& \leq C_2b\wt{\rho}M^2 + \frac{4b}{n}\biggl(1 + \frac{\sqrt{2}}{c_3^{2}}\wt{\rho}^{-d/4} + \frac{r^4n^{1/2}}{c_3^2\wt{\rho}^2}\biggr) \\ 
+& \leq \Bigl(C_2 + 4 + \frac{4\sqrt{2}}{c_3^{2}} + \frac{4}{c_3^2}\Bigr)bM^{2d/(4 + d)} n^{-4/(4 + d)}.
+\end{align*}
+The second inequality in the above is justified by Lemma~\ref{lem:variance_term_testing}, keeping in mind that $M \leq M_{\max}(d)$ implies that $\wt{\rho}^{-d/2} \leq n$.
+The third inequality follows from the upper bound on $r$ assumed in~\ref{asmp:ls_kernel_radius_testing} as well as the fact that $M \geq n^{-1/2}$.
+
+Next we lower bound the left hand side of the inequality \eqref{eqn:ls_fixed_graph_testing_critical_radius}---i.e.
 we lower bound the empirical norm $\|f_0\|_n^2$---using Lemma~\ref{lem:empirical_norm_sobolev}.
 Recall that by assumption, $M \leq M_{\max}(d)$.
-Therefore, taking $C \geq C_6$ in~\eqref{eqn:laplacian_smoothing_testing}, we have that~\eqref{eqn:empirical_norm_sobolev_1} is satisfied, so that it follows from~\eqref{eqn:empirical_norm_sobolev} that
-\begin{equation*}
-\|f\|_n^2 \geq \frac{\|f\|_{P}^2}{b} \geq \frac{p_{\min}}{b} \|f\|_{\Leb^2(\Xset)}^2 \geq CbM^{2d/(4 + d)}n^{-4/(4 + d)},
+Therefore, taking $C \geq C_6$ in \eqref{eqn:laplacian_smoothing_testing}, implies that the lower bound on $\|f\|_{\Leb^2(\mc{X})}$ in \eqref{eqn:empirical_norm_sobolev_1} is satisfied.
+As a result, it follows from \eqref{eqn:empirical_norm_sobolev} that
+\begin{equation*}
+\|f\|_n^2 \geq \frac{\mathbb{E}[\|f\|_n^2]}{b} \geq \frac{p_{\min}}{b} \|f\|_{\Leb^2(\Xset)}^2 \geq CbM^{2d/(4 + d)}n^{-4/(4 + d)},
 \end{equation*}
 with probability at least $1 - 5/b$.
-Taking $C \geq C_2 + 4 + (4\sqrt{2})/c_3^{d/4} + 4/c_3^2$ in~\eqref{eqn:laplacian_smoothing_testing} thus implies~\eqref{eqn:ls_fixed_graph_testing_critical_radius}, and we may therefore use Lemma~\ref{lem:ls_fixed_graph_testing} to upper bound the type II error the Laplacian smoothing test $\wh{\varphi}$.
-Observe that by~\eqref{eqn:neighborhood_eigenvalue} and the lower bound in Lemma~\ref{lem:variance_term_testing},
-\begin{equation*}
-\sum_{k = 1}^{n} \biggl(\frac{1}{\rho \lambda_k + 1}\biggr)^4 \geq 1 + \sum_{k = 2}^{n}\biggl(\frac{1}{C_3\wt{\rho}k^{2/d} + 1}\biggr)^4 \geq 1 + \frac{1}{16}\min\biggl\{\floor{\frac{\wt{\rho}^{-d/2}}{C_3^{d/2}}} - 1 \vee 0, n \biggr\} \geq \min\biggl\{\frac{\wt{\rho}^{-d/2}}{32 C_3^{d/2}}, \frac{n}{16} \biggr\}.
+Taking $C \geq C_2 + 4 + (4\sqrt{2})/c_3^{2} + 4/c_3^2$ in \eqref{eqn:laplacian_smoothing_testing} thus implies \eqref{eqn:ls_fixed_graph_testing_critical_radius}, and we may therefore use Lemma~\ref{lem:ls_fixed_graph_testing} to upper bound the type II error the Laplacian smoothing test $\wh{\varphi}$.
+Observe that by \eqref{eqn:neighborhood_eigenvalue} and the lower bound in Lemma~\ref{lem:variance_term_testing},
+\begin{equation*}
+\sum_{k = 1}^{n} \biggl(\frac{1}{\rho \lambda_k + 1}\biggr)^4 \geq 1 + \frac{1}{C_3^4}\sum_{k = 2}^{n}\biggl(\frac{1}{\wt{\rho}k^{2/d} + 1}\biggr)^4 \geq \frac{1}{32C_3^4}\wt{\rho}^{-d/2}.
 \end{equation*}
@@
 	\label{lem:variance_term_testing}
-	For any $t > 0$, when $d < 4$,
-	\begin{equation*}
-	\frac{1}{16}\min\Bigl\{\floor{t^{-d/2}} - 1 \vee 0,n\Bigr\} \leq \sum_{k = 2}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^4 \leq 2t^{-d/2}.
-	\end{equation*}
-\end{lemma}
+	Suppose $d \leq 4$. Then for any $t > 0$ such that $1 \leq t^{-d/2} \leq n$,
+	\begin{equation*}
+	\frac{1}{32}t^{-d/2} - 1 \leq \sum_{k = 2}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^4 \leq 2t^{-d/2}.
+	\end{equation*}
+\end{lemma}
+
 \paragraph{Proof of Lemma~\ref{lem:variance_term_estimation}.}
-We begin by proving the upper bounds. Observe that
-\begin{align*}
-\sum_{k = 1}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^2 & \leq t^{-d/2} + \frac{1}{t^2}\sum_{k = t^{-d/2} + 1}^{n} \frac{1}{k^{4/d}} \\
-& \leq t^{-d/2} + \frac{1}{t^2} \int_{t^{-d/2}}^{n} x^{-4/d} \,dx
-\end{align*}
-and the various upper bounds (for $d < 4$, $d = 4$, and $d > 4$) follow upon computing the integral.
-
-For the lower bound, we simply recognize that for each $k = 2,\ldots,n$ such that $k \leq \floor{t^{-d/2}}$, it holds that $1/(tk^{2/d} + 1)^2 \geq 1/4$, and there are at least $\min\Bigl\{\floor{t^{-d/2}} - 1 \vee 0,n\Bigr\}$ such values of $k$.
+We begin by proving the upper bounds. Treating the sum over $k$ as a Riemann sum of a non-increasing function, we have that
+\begin{align*}
+\sum_{k = 2}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^2 & \leq \int_{1}^{n} \biggl(\frac{1}{tx^{2/d} + 1}\biggr)^2 \,dx \leq t^{-d/2} + \int_{t^{-d/2}}^{n} \biggl(\frac{1}{tx^{2/d} + 1}\biggr)^2 \,dx \leq t^{-d/2} + \frac{1}{t^2} \int_{t^{-d/2}}^{n} x^{-4/d} \,dx.
+\end{align*}
+The various upper bounds (for $d < 4$, $d = 4$, and $d > 4$) then follow upon computing the integral.
+
+For the lower bound, we simply recognize that for each $k = 2,\ldots,n$ such that $k \leq \floor{t^{-d/2}}$, it holds that $1/(tk^{2/d} + 1)^2 \geq 1/4$, and there are at least $\min\Bigl\{\floor{t^{-d/2}} - 1,n - 1\Bigr\} > \frac{1}{2}t^{-d/2} - 1$ such values of $k$.
 
@@
 \begin{align*}
-\sum_{k = 1}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^4 & \leq t^{-d/2} + \frac{1}{t^4}\sum_{k = t^{-d/2} + 1}^{n} \frac{1}{k^{8/d}} \\
-& \leq t^{-d/2} + \frac{1}{t^4} \int_{t^{-d/2}}^{n} x^{-8/d} \,dx \\
-& \leq 2t^{-d/2}.
-\end{align*}
-
+\sum_{k = 1}^{n} \biggl(\frac{1}{tk^{2/d} + 1}\biggr)^4 \leq t^{-d/2} + \frac{1}{t^4}\sum_{k = t^{-d/2} + 1}^{n} \frac{1}{k^{8/d}} \leq t^{-d/2} + \frac{1}{t^4} \int_{t^{-d/2}}^{n} x^{-8/d} \,dx \leq 2t^{-d/2}.
+\end{align*}
 The lower bound follows from the same logic as we used to derive the lower bound in Lemma~\ref{lem:variance_term_estimation}.
 
-\section{Concentration Inequalities}
+\section{Concentration inequalities}
+
 \label{sec:concentration}

### 2021-03-03 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
Updated arXiv version of draft to be in-sync with AISTATS camera-ready submission.

@@
 \usepackage{fullpage}
+\usepackage{microtype}
+\usepackage{bm}
 
@@
 \begin{abstract}
-	In this paper we study the statistical properties of Laplacian smoothing, a graph-based approach to non-parametric regression.
-Under standard regularity conditions, we establish upper bounds on the error of the Laplacian smoothing estimator $\wh{f}$, and a goodness-of-fit test based on $\wh{f}$.
-These upper bounds match the minimax optimal estimation and testing rates of convergence over the first-order Sobolev class $H^1(\Xset)$, where $\Xset \subset \Rd$ and $1 \leq d < 4$; in the estimation problem, they are within a $\log n$ factor of optimal when $d = 4$.
-Additionally, we prove that Laplacian smoothing is manifold adaptive: if $\Xset$ is an $m$ dimensional manifold embedded in $\Reals^d$ with $m \ll d$, then the error rate of Laplacian smoothing methods depends only on the intrinsic dimension $m$ and not on the ambient dimension $d$.
+	In this paper we study the statistical properties of Laplacian smoothing, a graph-based approach to nonparametric regression.
+Under standard regularity conditions, we establish upper bounds on the error of the Laplacian smoothing estimator \smash{$\wh{f}$}, and a goodness-of-fit test also based on \smash{$\wh{f}$}.
+These upper bounds match the minimax optimal estimation and testing rates of convergence over the first-order Sobolev class $H^1(\Xset)$, for $\Xset \subseteq \Rd$ and $1 \leq d < 4$; in the estimation problem, for $d = 4$, they are optimal modulo a $\log n$ factor.
+Additionally, we prove that Laplacian smoothing is manifold-adaptive: if $\Xset \subseteq \Reals^d$ is an $m$-dimensional manifold with $m < d$, then the error rate of Laplacian smoothing (in either estimation or testing) depends only on $m$, in the same way it would if $\Xset$ were a full-dimensional set in $\Reals^m$.
 \end{abstract}
@@
 \section{INTRODUCTION}
-In the nonparametric regression problem, we observe samples $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to the signal plus noise model
+
+We adopt the standard setup in nonparametric regression, where we observe samples $(X_1,Y_1),\ldots,(X_n,Y_n)$ that are i.i.d.\ draws from the model 
 \begin{equation}
 \label{eqn:signal_plus_noise_model}
-Y_i = f_{0}(X_i) + \varepsilon_i,
+Y_i = f_0(X_i) + \varepsilon_i, \quad \epsilon_i \sim N(0,1), 
 \end{equation}
-with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise.
-Our goal is to perform statistical inference on the unknown regression function $f_0$, by which we mean either \emph{estimating} $f_0$ or more simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present.
+where $\varepsilon_i$ is independent of $X_i$.
+Our goal is to perform statistical inference on the unknown regression function $f_0$, by which we mean either \emph{estimating} $f_0$ or more simply \emph{testing} whether $f_0 = 0$, i.e., whether there is any signal present.
 
-The Laplacian smoothing estimator $\wh{f}$ \citep{smola2003} is a penalized least squares estimator, defined on a graph.
-Letting $G = \bigl(\{1,\ldots,n\},W\bigr)$ be a weighted graph over vertices $\{1,\ldots,n\}$ which we associate with the sample points $\{X_1,\ldots,X_n\}$, the estimator $\wh{f}$ is given by
+Laplacian smoothing \citep{smola2003} is a penalized least squares estimator, defined over a graph.
+Letting $G = (V,W)$ be a weighted undirected graph with vertices $V=\{1,\ldots,n\}$, associated with $\{X_1,\ldots,X_n\}$, the Laplacian smoothing estimator \smash{$\wh{f}$} is given by
 \begin{equation}
 \label{eqn:laplacian_smoothing}
-\wh{f} :=  \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap f \biggr\}.
+\wh{f} =  \argmin_{f \in \Reals^n} \; \sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^\top \Lap f. 
 \end{equation}
-Here $\Lap$ is the graph Laplacian matrix (defined in Section~\ref{sec:problem_setup_and_background}), $G$ is typically a geometric graph (such as a $k$NN or neighborhood graph) and the penalty
+Here $\Lap$ is the graph Laplacian matrix (to be defined in Section~\ref{sec:problem_setup_and_background}), $G$ is typically a geometric graph (such as a $k$-nearest-neighbor or neighborhood graph), $\rho \geq 0$ is a tuning parameter, and the penalty
 \begin{equation*}
-f^T \Lap f = \frac{1}{2} \sum_{i,j = 1}^{n} W_{ij}\bigl(f_i - f_j\bigr)^2
+f^\top \Lap f = \frac{1}{2} \sum_{i,j = 1}^{n} W_{ij}(f_i - f_j)^2
 \end{equation*}
-thus encourages $\wh{f}_i \approx \wh{f}_j$ when $X_i \approx X_j$. Assuming~\eqref{eqn:laplacian_smoothing} is a reasonable estimator of $f_0$, the statistic
+encourages \smash{$\wh{f}_i \approx \wh{f}_j$} when $X_i \approx X_j$. Assuming \eqref{eqn:laplacian_smoothing} is a reasonable estimator of $f_0$, the statistic
 \begin{equation}
 \label{eqn:laplacian_smoothing_test}
-\wh{T} := \frac{1}{n}\bigl\|\wh{f}\bigr\|_2^2 
+\wh{T} = \frac{1}{n} \| \wh{f} \|_2^2 
 \end{equation}
-is in turn a reasonable way to decide if $f_0 = 0$.
+would in turn be a reasonable way to test $f_0 = 0$.
 
-Of course there exist many methods for nonparametric regression \citep{gyorfi2006,wasserman2006,tsybakov2008_book}, but Laplacian smoothing has its own set of advantages. For instance:
+Of course there are many methods for nonparametric regression (see, e.g., \citet{gyorfi2006,wasserman2006,tsybakov2008_book}), but Laplacian smoothing has its own set of advantages. For instance:
 \begin{itemize}
 	\item \emph{Computational ease.} Laplacian smoothing is fast, easy, and stable to compute.
-The estimate $\wh{f}$ can be computed by solving a symmetric, diagonally dominant linear system.
-There are known extremely fast and reliable solvers for exactly this problem (see e.g.
-the seminal papers of \citet{spielman2011,spielman2013,spielman2014}, or the monograph by~\citet{vishnoi2012} and references therein).
-	\item \emph{Generality.} Laplacian smoothing is well defined whenever one can associate a graph with some observed responses.
-This generality lends itself to many different data modalities: for instance text and image classification, to name only two \citep{kondor2002, belkin03a,belkin2006}.
-	\item \emph{Weak supervision.} Although we study Laplacian smoothing in the supervised learning setup~\eqref{eqn:signal_plus_noise_model}, it can be easily adapted to the semi-supervised or unsupervised settings \citep{nadler09,slepcev17,calder2019b,dunlop2020}.
+The estimate \smash{$\wh{f}$} can be computed by solving a symmetric diagonally dominant linear system.
+There are by now various nearly-linear-time solvers for this problem (see e.g., the seminal papers of \citet{spielman2011,spielman2013,spielman2014}, or the overview by \citet{vishnoi2012} and references therein).
+	\item \emph{Generality.} Laplacian smoothing is well-defined whenever one can associate a graph with observed responses.
+This generality lends itself to many different data modalities, e.g., text and image classification, as in \citet{kondor2002,belkin03a,belkin2006}.
+	\item \emph{Weak supervision.} Although we study Laplacian smoothing in the supervised problem setting \eqref{eqn:signal_plus_noise_model}, the method can be adapted to the semi-supervised or unsupervised settings, as in \citet{zhu2003semisupervised,zhou2005learning,nadler09}.
+%,slepcev17,calder2019b,dunlop2020}.
 \end{itemize}
 
-For these reasons, a body of work has emerged analyzing the statistical properties of Laplacian smoothing, and graph-based methods more generally.
-Roughly speaking, these works can be divided into two categories, based on the perspective they adopt.
+For these reasons, a body of work has emerged that analyzes the statistical properties of Laplacian smoothing, and graph-based methods more generally.
+Roughly speaking, this work can be divided into two categories, based on the perspective they adopt.
 \begin{itemize}
-	\item In the \emph{fixed design perspective}, the design points $\{X_1,\ldots,X_n\}$ and the graph $G$ are treated as fixed, and inference is carried out on the evaluations $\bigl(f_0(X_1),\ldots,f_0(X_n)\bigr) \in \Reals^n$.\footnote{To be clear, while in this setting $G$ may be a geometric graph, in principle it could be any graph over $\{1,\ldots,n\}$.} In this context, tight upper bounds have been derived on the error of graph-based estimators \citep{wang2016, sadhanala16,sadhanala17,kirichenko2017,kirichenko2018} and tests \citep{sharpnack2013,sharpnack2013b,sharpnack2015}, which certify that such procedures are \emph{optimal} over ``function'' classes $\mc{F} \subseteq \Reals^n$.
-The downside of this work is that it may be unnatural to assume a priori that the graph $G$ is fixed, or that the evaluations $\bigl(f_0(X_1),\ldots,f_0(X_n)\bigr)$ belong to some discrete ``function'' class.
-	\item In the arguably more natural \emph{random design perspective},  one assumes the design points $X_1,\ldots,X_n$ are sampled independently from some distribution $P$ supported on a domain $\Xset \subset \Rd$.
-Inference is drawn on the regression function $f_0: \Xset \to \Reals$, which is typically assumed to be smooth in some continuum sense, e.g.
-it possesses a first derivative bounded in $L^{\infty}$ (H\"{o}lder) or $\Leb^2$ (Sobolev) norm.
-To conduct graph-based inference, the user first builds a neighborhood graph over the random design points---so that $W_{ij}$ is large iff $X_i$ and $X_j$ are close in say Euclidean distance---and then solves e.g.~\eqref{eqn:laplacian_smoothing} or~\eqref{eqn:laplacian_smoothing_test}.
-In this context, various graph-based procedures have been shown to be \emph{consistent}: as $n \to \infty$, they converge to a \emph{continuum} limit (see \citet{belkin07,trillos2018,vonluxburg2008} among others).
-However, until recently such statements were not accompanied by error rates, and such error rates as have been proved \citep{lee2016,trillos2020} are not optimal over continuum function spaces, such as H\"{o}lder or Sobolev classes.
+	\item \emph{Fixed design perspective.} Here one treats the design points $X_1,\ldots,X_n$ and the graph $G$ as fixed, and carries out inference on $f_0(X_i)$, $i=1,\ldots,n$.
+In this problem setting, tight upper bounds have been derived on the error of various graph-based methods (e.g., \citet{wang2016,hutter2016,sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}) and tests (e.g., \citet{sharpnack2010identifying,sharpnack2013b,sharpnack2013,sharpnack2015}), which certify that such procedures are \emph{optimal} over ``function'' classes (in quotes because these classes really model the $n$-dimensional vector of evaluations).
+The upside of this work is its generality: in this setting $G$ need not be a geometric graph, but in principle it could be any graph over $V=\{1,\ldots,n\}$.
+The downside is that, in the context of nonparametric regression, it is not as natural to think of the evaluations of $f_0$ as exhibiting smoothness over some fixed pre-defined graph $G$, and more natural to speak of the smoothness of the function $f_0$ itself.
+	\item \emph{Random design perspective.} Here one treats the design points $X_1,\ldots,X_n$ as independent samples from some distribution $P$ supported on a domain $\Xset \subseteq \Rd$.
+Inference is drawn on the regression function $f_0: \Xset \to \Reals$, which is typically assumed to be smooth in some \emph{continuum} sense, e.g., it possesses a first derivative bounded in $L^{\infty}$ (H\"{o}lder) or $\Leb^2$ (Sobolev) norm.
+To conduct graph-based inference, the user first builds a neighborhood graph over the random design points---so that $W_{ij}$ is large when $X_i$ and $X_j$ are close in (say) Euclidean distance---and then computes e.g., \eqref{eqn:laplacian_smoothing} or \eqref{eqn:laplacian_smoothing_test}.
+In this context, various graph-based procedures have been shown to be \emph{consistent}: as $n \to \infty$, they converge to a continuum limit (see \citet{belkin07,vonluxburg2008,trillos2018} among others).
+However, until recently such statements were not accompanied by error rates, and even so, such error rates as have been proved \citep{lee2016,trillos2020} are not optimal over continuum function spaces, such as H\"{o}lder or Sobolev classes.
 \end{itemize}
-Hereafter, we will adopt the random design perspective, and seek to answer the following question:
-\begin{quote}{When $f_0$ is assumed to smooth in a continuum sense, does Laplacian smoothing achieve optimal performance for estimation and goodness-of-fit testing?} 
+The random design perspective bears a more natural connection with nonparametric regression (the focus in this paper), as it allows us to formulate smoothness based on $f_0$ itself (how it behaves as a continuum function, and not just its evaluations at the design points).
+In this paper, we will adopt the random design perspective, and seek to answer the following question:
+\begin{quote}{When we assume the regression function $f_0$ to be smooth in a continuum sense, does Laplacian smoothing achieve optimal performance for estimation and goodness-of-fit testing?} 
 \end{quote}
-On the one hand, this is no small question---rather, it is arguably the fundamental question of nonparametric regression, and without an answer one cannot fully compare the statistical properties of Laplacian smoothing to its competitors.
-On the other hand, it seems difficult to answer: there is a fundamental gap between the \emph{discrete} smoothness imposed by the penalty $f^T \Lap f$ in~\eqref{eqn:laplacian_smoothing} and the \emph{continuum} smoothness assumed on $f_0$, and to obtain sharp upper bounds we will need to bridge this gap in a tight sense.
+This is no small question---arguably, it is \emph{the} central question of nonparametric regression---and without an answer one cannot fully compare the statistical properties of Laplacian smoothing to alternative methods.
+It also seems difficult to answer: as we discuss next, there is a fundamental gap between the \emph{discrete} smoothness imposed by the penalty $f^\top \Lap f$ in problem \eqref{eqn:laplacian_smoothing} and the \emph{continuum} smoothness assumed on $f_0$, and in order to obtain sharp upper bounds we will need to bridge this gap in a suitable sense.
 
@@
 
-\paragraph{Advantages of the discrete approach.} In light of this last point, it is worth asking whether there is any \emph{statistical} advantage to solving a discrete problem such as~\eqref{eqn:laplacian_smoothing} (setting aside computational considerations for the moment).
+\paragraph{Advantages of the Discrete Approach.}
+
+In light of this last point, it is worth asking whether there is any \emph{statistical} advantage to solving a discrete problem such as \eqref{eqn:laplacian_smoothing} (setting aside computational considerations for the moment).
 After all, we could have instead solved the following variational problem:
@@
 \label{eqn:thin_plate_spline}
-\wt{f} := \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \cdot \int_{\Xset} \|\nabla f(x)\|^2 \,dx \biggr\},
+\wt{f} = \argmin_{f : \Xset \to \Reals} \; \sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \hspace{-2pt} \int_{\Xset} \|\nabla f(x)\|_2^2 \,dx,
 \end{equation}
-with the minimum computed over all functions $f$ for which the criterion is finite--or used the statistic
+where the optimization is performed over all functions for which the criterion is well-defined and finite (continuous functions $f$ that have a weak derivative $\nabla f$ in $L^2(\Xset)$).
+Analogously, for testing, we could use:
 \begin{equation}
 \label{eqn:thin_plate_spline_test}
-\wt{T} := \|\wt{f}\|_n^2.
+\wt{T} = \| \wt{f} \|_n^2 = \frac{1}{n} \sum_{i=1}^n \wt{f}(X_i)^2.
 \end{equation}
-The penalty in~\eqref{eqn:thin_plate_spline} leverages the assumption that $f_0$ possesses a bounded derivative in a seemingly natural manner.
-Indeed, for this reason the estimator $\wt{f}$ (and to a lesser extent, the statistic $\wt{T}$) are well known--- in the univariate case, $\wt{f}$ is simply a \emph{smoothing spline}, and when $d > 1$ it is an example of a \emph{thin-plate spline}---and their statistical properties are well-understood \citep{vandergeer2000, liu2019}.
-Moreover, there exist strong connections between smoothing/thin-plate splines and Laplacian smoothing: as we show later on in~\eqref{eqn:seminorm_consistency}, in a certain sense the Laplacian smoothing problem~\eqref{eqn:laplacian_smoothing} can be seen as a discrete and noisy approximation to~\eqref{eqn:thin_plate_spline}.
-At first blush, this suggests that Laplacian smoothing should at best inherit the statistical properties of smoothing/thin-plate splines, and at worst may have meaningfully larger error.
+The penalty term in \eqref{eqn:thin_plate_spline} leverages the assumption that $f_0$ has a smooth derivative in a seemingly natural way.
+Indeed, the estimator \smash{$\wt{f}$} and statistic \smash{$\wt{T}$} are well-known: for $d = 1$, \smash{$\wt{f}$} is the familiar \emph{smoothing spline}, and for $d>1$, it is type of a \emph{thin-plate spline}.
+The statistical properties of smoothing and thin-plate splines are well-understood \citep{vandergeer2000,liu2019}.
+As we discuss later, the Laplacian smoothing problem \eqref{eqn:laplacian_smoothing} can be viewed as a discrete and noisy approximation to \eqref{eqn:thin_plate_spline}.
+At first blush, this suggests that Laplacian smoothing should at best inherit the statistical properties of \eqref{eqn:thin_plate_spline}, and at worst may have meaningfully larger error.
 
-However, as we shall see the actual story is quite different: remarkably, Laplacian smoothing enjoys optimality properties in cases where the thin-plate spline problem~\eqref{eqn:thin_plate_spline} is not even well-posed.
-Tables~\ref{tbl:estimation_rates} and~\ref{tbl:testing_rates} summarize the gap.
+However, as we shall see the actual story is quite different: remarkably, Laplacian smoothing enjoys optimality properties even in settings where the thin-plate spline estimator \eqref{eqn:thin_plate_spline} is not well-posed (to be explained shortly); Tables~\ref{tbl:estimation_rates} and~\ref{tbl:testing_rates} summarize.
 As we establish in Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing_manifold}, when computed over an appropriately formed neighborhood graph, Laplacian smoothing estimators and tests are minimax optimal over first-order \emph{continuum} Sobolev balls.
-This holds true either when $\Xset \subset \Rd$ is a full-dimensional domain and $d = 1,2$ or $3$, or when $\Xset$ is a submanifold of $\Rd$ of intrinsic dimension $m = 1,2,$ or $3$.
-Additionally, the estimator $\wh{f}$ is nearly minimax optimal (to within a $(\log n)^{1/3}$ factor) when $d$ (or $m$) = $4$.
-By contrast, smoothing splines are optimal only when $d$ (or $m$) = $1$.
-In higher dimensions, technically speaking the problem~\eqref{eqn:thin_plate_spline} is not even well-posed, and any practical implementation of the estimator $\wt{f}$ will be inconsistent \citep{green93}.
-Hence, our results favorably distinguish Laplacian smoothing methods from their natural variational alternatives.
+This holds true either when $\Xset \subseteq \Rd$ is a full-dimensional domain and $d = 1,2,$ or $3$, or when $\Xset$ is a manifold embedded in $\Rd$ of intrinsic dimension $m = 1,2,$ or $3$.
+Additionally, the estimator \smash{$\wh{f}$} is nearly minimax optimal (to within a \smash{$(\log n)^{1/3}$} factor) when $d=4$ (or $m=4$ in the manifold case).
+
+By contrast, smoothing splines are optimal only when $d=1$.
+When $d>1$, thin-plate spline estimator \eqref{eqn:thin_plate_spline} is not even well-posed, in the following sense: for any $(X_1,Y_1),\ldots,(X_n,Y_n)$ and any $\delta>0$, there exists (e.g., \citet{green93} give a construction using ``bump'' functions) a differentiable function $f$ such that $f(X_i) = Y_i$, $i=1,\ldots,n$, and
+\begin{equation*}
+\int_{\Xset} \|\nabla f(x)\|_2^2 \leq \delta.
+\end{equation*}
+In other words, $f$ achieves perfect (zero) data loss and arbitrarily small penalty in the problem \eqref{eqn:thin_plate_spline}.
+This will clearly not lead to a consistent estimator of $f_0$ across the design points (as it always yields $Y_i$ at each $X_i$).
+In this light, our results when $d>1$ favorably distinguish Laplacian smoothing from its natural variational analog.
+
 \begin{table}
@@
 			\hline
-			$d = 1$ & ${\bf n^{-2/3}}$ & \textcolor{red}{${\bf n^{-2/3}}$} \\
-			$d = 2,3$ & ${\bf n^{-2/(2 + d)}}$ & \textcolor{red}{$1$} \\
-			$d = 4$ & ${\bf n^{-1/3}} (\log n)^{1/3}$ & \textcolor{red}{$1$} \\
+			$d = 1$ & $\bm{n^{-2/3}}$ & \textcolor{red}{$\bm{n^{-2/3}}$} \\
+			$d = 2,3$ & $\bm{n^{-2/(2 + d)}}$ & \textcolor{red}{$1$} \\
+			$d = 4$ & $\bm{n^{-1/3}} (\log n)^{1/3}$ & \textcolor{red}{$1$} \\
 			$d \geq 5$  & $(\log n/n)^{4/(3d)}$ &\textcolor{red}{$1$} \\
@@
 	\caption{Summary of estimation rates over first-order Sobolev balls.
-Rates in black are new results of this paper, and bolded rates are minimax optimal.
-Although we suppress it for conciseness, in all cases the dependence of the error rate on the radius of the Sobolev ball is also optimal.
-For thin-plate splines, ``\textcolor{red}{$1$}'' indicates that the estimator is inconsistent.
-When $\Xset$ is an $m$ dimensional manifold embedded in $\Rd$, all rates hold with $d$ replaced by $m$, without any change in the method in the case of Laplacian smoothing.}
+Black font marks new results from this paper, red font marks previously-known results; bold font marks minimax optimal rates.
+Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius of the Sobolev ball is also optimal.
+We use ``$1$'' to indicate inconsistency (error not converging to 0).
+Furthermore, when $\Xset$ is an $m$-dimensional manifold embedded in $\Rd$, all Laplacian smoothing results hold with $d$ replaced by $m$, without any change to the method itself.}
 	\label{tbl:estimation_rates}
@@
 			\hline
-			$d = 1$ & ${\bf n^{-4/5}}$ & \textcolor{red}{${\bf n^{-4/5}}$} \\
-			$d = 2,3$ & ${\bf n^{-4/(4 + d)}}$ & \textcolor{red}{$n^{-1/2}$} \\
-			$d \geq 4$\footnotemark & ${ \bf n^{-1/2}}$ & \textcolor{red}{${ \bf n^{-1/2}}$}
+			$d = 1$ & $\bm{n^{-4/5}}$ & \textcolor{red}{$\bm{n^{-4/5}}$} \\
+			$d = 2,3$ & $\bm{n^{-4/(4 + d)}}$ & \textcolor{red}{$n^{-1/2}$} \\
+			$d \geq 4$ & $\bm{n^{-1/2}}$ & \textcolor{red}{$\bm{n^{-1/2}}$}
 		\end{tabular}
 	\end{center}
-	\caption{Summary of testing rates over fixed radius first-order Sobolev balls; black, red, and bold font used as in Table~\ref{tbl:estimation_rates}.
-Rates for $d \geq 4$ assume $f_0 \in \Leb^4(\Xset,M)$.
-When $\Xset$ is an $m$ dimensional manifold embedded in $\Rd$, all rates hold with $d$ replaced by $m$.}
+	\caption{Summary of testing rates over first-order Sobolev balls; black, red, and bold fonts are used as in Table~\ref{tbl:estimation_rates}.
+Rates for $d \geq 4$ assume that $f_0 \in \Leb^4(\Xset,M)$.
+When $\Xset$ is an $m$-dimensional manifold embedded in $\Rd$, all rates hold with $d$ replaced by $m$.}
 	\label{tbl:testing_rates}
 \end{table}
-\footnotetext{Assuming $f_0 \in \Leb^4(\Xset,M)$.}
-\paragraph{Future extensions.}
-To be clear, these results are not the end of the story.
-For one thing, the estimator $\wh{f}$ is defined only at the design points $\{X_1,\ldots,X_n\}$.
-As such, we measure loss using the in-sample mean squared error
+%\footnotetext{Assuming $f_0 \in \Leb^4(\Xset,M)$.}
+
+\paragraph{Future Directions.}
+
+To be clear, there is still much left to be investigated.
+For one, the Laplacian smoothing estimator \smash{$\wh{f}$} is only defined at $X_1,\ldots,X_n$.
+In this work we study its in-sample mean squared error
+\begin{equation}
+\label{eqn:empirical_norm_error}
+\bigl\|\wh{f} - f_0\bigr\|_n^2 := \frac{1}{n}\sum_{i = 1}^{n} \Bigl(\wh{f}_i - f_0(X_i)\Bigr)^2.
+\end{equation}
+In Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we discuss how to extend \smash{$\wh{f}$} to a function over all $\Xset$, in such a way that the out-of-sample mean squared error \smash{$\|\wh{f} - f_0\|_{\Leb^2(\Xset)}^2$} should remain small, but leave a formal analysis to future work.
+
+In a different direction, problem \eqref{eqn:thin_plate_spline} is only a special, first-order case of thin-plate splines. In general, the $k$th order thin-plate spline estimator is defined as
 \begin{equation*}
-\|\wh{f} - f_0\|_n^2 := \frac{1}{n}\sum_{i = 1}^{n} \Bigl(\wh{f}_i - f_0(X_i)\Bigr)^2.
+\wt{f} = \argmin_{f : \Xset \to \Rd} \; \sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \hspace{-2pt} \sum_{|\alpha| = k} \int_{\Xset} \bigl(D^\alpha f(x)\bigr)^2 \,dx,
 \end{equation*}
-In Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we briefly discuss how one could extend $\wh{f}$ to a function over all $\Xset$, in such a way that the out-of-sample mean square error $\|\wh{f} - f_0\|_{\Leb^2(\Xset)}$ remains small, but leave a formal analysis of the problem to future work.
+where for each multi-index $\alpha=(\alpha_1,\ldots,\alpha_d)$ we write $D^\alpha f(x) = \partial^kf/\partial x_{1}^{\alpha_1} \cdots \partial x_{d}^{\alpha_d}$.
+This problem is in general well-posed whenever $2k > d$.
+In this regime, assuming that the $k$th order partial derivatives $D^\alpha f_0$ are all $\Leb^2(\Xset)$ bounded, the degree $k$ thin-plate spline has error on the order of \smash{$n^{-2k/(2k + d)}$} \citep{vandergeer2000}, which is minimax rate-optimal for such functions.
+Of course, assuming $f_0$ has $k$ bounded derivatives for some $2k > d$ is a very strong condition, but at present we do not know if (adaptations of) Laplacian smoothing on neighborhood graphs achieve these rates.
 
-In a different direction, the problem~\eqref{eqn:thin_plate_spline} is only a special, first-order case of thin-plate splines. In general, the degree-$k$ thin plate spline is defined as a minimizer of
+\paragraph{Notation.}
+
+For an integer $p \geq 1$, we use $\Leb^p(\Xset)$ for the set of functions $f$ such that 
 \begin{equation*}
-\sum_{i = 1}^{n} (Y_i - f(X_i))^2 + \rho \cdot \sum_{|m| = k} \int_{\Xset} \bigl(D^mf(x)\bigr)^2 \,dx
+\norm{f}_{\Leb^p(\Xset)}^p := \int_{\Xset} |f(x)|^p \,dx < \infty,
 \end{equation*}
-where for each multiindex $m$ we write $D^m$ for the partial derivative $D^mf(x):= \partial^kf/\partial x_{1}^{m_1}\ldots \partial x_{d}^{m_d}$.
-Assuming the $k$th-order partial derivatives $D^mf_0$ are all bounded in $\Leb^2(\Xset)$ norm, the degree-$k$ thin plate spline has error on the order of $n^{-2k/(2k + d)}$ \citep{vandergeer2000}, which is the minimax optimal rate for such functions.
-Of course, assuming $f_0$ possesses $k$ bounded derivatives for some $2k > d$ is a rather strong assumption which may not be warranted, but at present we do not know whether (adaptations of) Laplacian smoothing on neighborhood graphs can achieve these faster rates.
-
-% \paragraph{Notation.}
-% AJG: Need to make sure these are all defined somewhere
-% For an integer $p > 0$, we use $\Leb^p(\Xset)$ to refer to the set of functions $f$ for which the $p$th moment $\norm{f}_{\Leb^p(\Xset)}^p := \int_{\Xset} |f(x)|^p \,dx < \infty$, and $C^p(\Xset)$ to refer to those function which are $p$-times continuously differentiable with bounded $p$th derivative.
-Finally, we use $C,C_1,\ldots$ and $c,c_2,\ldots$ as constants which are independent of $n$ and $f_0$, and for convenience, we will also use asymptotic notation: $a_n \lesssim b_n$ means that $a_n \leq Cb_n$ for some $C > 0$, and $a_n \asymp b_n$ means that $a_n \lesssim b_n$ and $b_n \lesssim a_n$.
+and $C^p(\Xset)$ for the set of functions that are $p$ times continuously differentiable.
+For sequences $a_n,b_n$, we write $a_n \lesssim b_n$ to mean $a_n \leq Cb_n$ for a constant $C>0$ and large enough $n$, and $a_n \asymp b_n$ to mean $a_n \lesssim b_n$ and $b_n \lesssim a_n$.
+Lastly, we use $a \wedge b = \max\{a,b\}$.
 
@@
 \label{sec:problem_setup_and_background}
-Before we formally state our main results in Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we define neighborhood graph Laplacians, and review the known minimax rates over first-order Sobolev spaces.
 
-\paragraph{Neighborhood graph Laplacians.}
-As mentioned, in the graph-based approach to nonparametric regression we first build a neighborhood graph $G_{n,r} = ([n],W)$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense.
-The $n \times n$ weight matrix $W = (W_{ij})$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $W_{ij}$ are given by
+Before we present our main results in Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we define neighborhood graph Laplacians, and review known minimax rates over first-order Sobolev spaces.
+
+\paragraph{Neighborhood Graph Laplacians.}
+
+In the graph-based approach to nonparametric regression, we first build a neighborhood graph \smash{$G_{n,r} = (V,W)$}, for $V=\{1,\ldots,n\}$, to capture the geometry of $P$ (the design distribution) and $\Xset$ (the domain) in a suitable sense.
+The $n \times n$ weight matrix $W = (W_{ij})$ encodes proximity between pairs of design points; for a kernel function $K: [0,\infty) \to \Reals$ and radius $r > 0$, we have
 \begin{equation*}
 \label{eqn:neighborhood_graph}
-W_{ij} = K\Biggl(\frac{\norm{X_i - X_j}}{r}\Biggr),
+W_{ij} = K\Biggl(\frac{\|X_i - X_j\|_2}{r}\Biggr),
 \end{equation*}
-where $\norm{\cdot}$ is the Euclidean norm in $\Rd$.
-Then the degree matrix $D$ is the $n \times n$ diagonal matrix with entries $D_{ii} = \sum_{j = 1}^{n}W_{ij}$, and the graph Laplacian can be written as
+with $\|\cdot\|_2$ denoting the $\ell_2$ norm on $\Rd$.
+Defining $D$ as the $n \times n$ diagonal matrix with entries \smash{$D_{ii} = \sum_{j = 1}^{n} W_{ij}$}, the graph Laplacian can then be written as
 \begin{equation}
 \label{eqn:graph_Laplacian}
-\Lap_{n,r} = D - W;
+\Lap_{n,r} = D - W.
 \end{equation}
-we add the subscripts $n$ and $r$ to $\Lap$ and $G$ to emphasize that they depend on the random design points $X_1,\ldots,X_n$ and the radius $r$.
-To avoid any confusion: henceforth when speak of $\wh{f}$ and $\wh{T}$, we will always assume they are fit using the neighborhood graph Laplacian $L = L_{n,r}$.
+We add subscripts $n,r$ to $\Lap$ and $G$ to emphasize that they depend on the design points $X_1,\ldots,X_n$ and the radius $r$.
+To avoid any confusion: henceforth when we speak of \smash{$\wh{f}$} and \smash{$\wh{T}$}, we will always assume that they are fit using the neighborhood graph Laplacian \smash{$L = L_{n,r}$}.
+Writing its eigendecomposition as \smash{$L = \sum_{k = 1}^{n} \lambda_k v_k v_k^\top$}, we always assume, by convention, ordered eigenvalues $0 = \lambda_1 \leq \cdots \leq \lambda_n$, and unit-norm eigenvectors.
+%we will make reference to these objects later on.
 
-Note that $L$ is a symmetric, positive semi-definite matrix, and so admits an eigendecomposition $L = \sum_{k = 1}^{n} \lambda_k v_k v_k^T$.
-We order the eigenvalues $0 = \lambda_1 \leq \lambda_2 \leq \cdots \leq \lambda_n$, and always consider unit-norm eigenvectors $\|v_k\|_2^2 = 1$; we will make reference to these objects later on.
+\paragraph{Sobolev Spaces.}
 
-\paragraph{Sobolev spaces.}
-Now that we've defined the neighborhood graph Laplacian $\Lap_{n,r}$, we step away from graph-based methods for a moment, to briefly recall some classical results regarding minimax rates over Sobolev classes.
-
-We say that a function $f \in \Leb^2(\Xset)$ belongs to the \emph{first-order Sobolev space} $H^1(\Xset)$ iff for each $j = 1,\ldots,d$, the weak partial derivative $D^jf$ exists and belong to $\Leb^2(\Xset)$.
-For such functions $f \in H^1(\Xset)$, the Sobolev seminorm $\seminorm{f}_{H^{1}(\Xset)}$ is the average size of the gradient;
+We step away from graph-based methods for a moment, to briefly recall some classical results regarding minimax rates over Sobolev classes.
+We say that a function $f \in \Leb^2(\Xset)$ belongs to the \emph{first-order Sobolev space} $H^1(\Xset)$ if, for each $j = 1,\ldots,d$, the weak partial derivative $D^j f$ exists and belongs to $\Leb^2(\Xset)$.
+For such functions $f \in H^1(\Xset)$, the Sobolev seminorm \smash{$\seminorm{f}_{H^{1}(\Xset)}$} is the average size of the gradient $\nabla f = (D^1 f, \ldots, D^d f)$,
 \begin{equation*}
-\seminorm{f}_{H^1(\Xset)}^2 = \int_{\Xset} \bigl\|\nabla f(x)\bigr\|^2 \,dx
+\seminorm{f}_{H^1(\Xset)}^2 := \int_{\Xset} \bigl\|\nabla f(x)\bigr\|_2^2 \,dx,
 \end{equation*}
-with corresponding Sobolev norm $\norm{f}_{H^1(\Xset)} := \norm{f}_{\Leb^2(\Xset)} + |f|_{H^1(\Xset)}$. The Sobolev balls $H^1(\Xset,M)$ for $M > 0$ are then defined as
+with corresponding Sobolev norm 
+\begin{equation*}
+\norm{f}_{H^1(\Xset)} := \norm{f}_{\Leb^2(\Xset)} + |f|_{H^1(\Xset)}.
+\end{equation*}
+The Sobolev ball $H^1(\Xset,M)$ for $M > 0$ is
 \begin{equation*}
@@
 \end{equation*}
-For further details regarding Sobolev spaces see e.g.~\citet{evans10,leoni2017}.
+For further details regarding Sobolev spaces see, e.g., \citet{evans10,leoni2017}.
 
-\paragraph{Minimax rates.}
-To carry out a minimax analysis of regression in Sobolev spaces, one must impose some regularity conditions on the design distribution $P$. We shall assume the following.
+\paragraph{Minimax Rates.}
+
+To carry out a minimax analysis of regression in Sobolev spaces, one must impose regularity conditions on the design distribution $P$. We shall assume the following.
 \begin{enumerate}[label=(P\arabic*)]
@@
 	\label{asmp:domain}
-	$P$ is supported on a domain $\Xset \subset \Rd$, an open, connected set with Lipschitz boundary.
+	$P$ is supported on a domain $\Xset \subseteq \Rd$, which is an open, connected set with Lipschitz boundary.
 	\item
-	\label{asmp:bounded_lipschitz_density} 
-	$P$ admits a density $p$ bounded away from $0$ and $\infty$, i.e.
+	\label{asmp:density} 
+	$P$ admits a density $p$ such that
 	\begin{equation*}
-	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty,~~\textrm{for all $x \in \Xset$.}
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty, ~~\textrm{for all $x \in \Xset$}.
 	\end{equation*}
-	Additionally, the restriction of $p$ to $\Xset$ is Lipschitz, with Lipschitz constant $L_p$.
+	Additionally, $p$ is Lipschitz on $\Xset$, with Lipschitz constant $L_p$.
 \end{enumerate}
 
-Under~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density}, the minimax estimation rate over Sobolev balls of radius $M \geq n^{-1/2}$ is (see e.g. \citet{tsybakov2008_book}),
+Under conditions~\ref{asmp:domain}, \ref{asmp:density}, the minimax estimation rate over a Sobolev ball of radius $M \geq n^{-1/2}$ is (e.g., \citet{tsybakov2008_book}):
 \begin{equation}
 \label{eqn:sobolev_space_estimation_minimax_rate}
-\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{2d/(2 + d)}n^{-2/(2 + d)}.
+\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{2d/(2 + d)} n^{-2/(2 + d)}.
 \end{equation}
-(Here and throughout, we assume $M \geq n^{-1/2}$, since otherwise the minimax rate is simply the parametric rate $n^{-1}$.)
+% RJT: This is a bit of a weird reference, because (as per my understanding) Tsybakov only proves the rate is tight for Holder classes, which are contained in the Sobolev class.
+So technically from Tsybakov alone we only know \gtrsim, not \asymp for this one.
+It's fine, I wouldn't change anything because it would involve explaining a lot more, I'm just pointing this out.
+(Throughout we assume $M \geq n^{-1/2}$, as otherwise the minimax rate is just the parametric rate $n^{-1}$.)
 
 As nonparametric hypothesis testing is (comparatively) less familiar than nonparametric estimation, we briefly summarize the main idea before stating the optimal error rate.
-In the goodness-of-fit testing problem, we ask for a test function---formally, a Borel measurable function $\phi$ that takes values in $\{0,1\}$--- which can distinguish between the hypotheses
+In the goodness-of-fit testing problem, we ask for a test function---formally, a Borel measurable function $\phi$ taking values in $\{0,1\}$---which can distinguish between the hypotheses
 \begin{equation}
@@
 Typically, the null hypothesis $f_0 = f_0^{\star} \in \mc{F}$ reflects the absence of interesting structure, and $\mc{F} \setminus  \{f_0^{\star}\}$ is a set of smooth departures from this null.
-To fix ideas, as in \citet{ingster2009} we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\star} = 0$ and $\mc{F} = H^1(\Xset,M)$ is a first-order Sobolev ball; this is without loss of generality.
+In this paper, as in \citet{ingster2009}, we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\star} = 0$ and $\mc{F} = H^1(\Xset,M)$ is a first-order Sobolev ball.
 
@@
 \begin{equation*}
-\epsilon(\mc{F}) := \inf\Biggl\{\epsilon > 0: \inf_{\phi} \biggl[ \sup_{f_0 \in \mc{F}_{\epsilon}} \Ebb_{f_0}[1 - \phi]\biggr] \leq \alpha\Biggr\}
+\epsilon(\mc{F}) := \inf\biggl\{\epsilon > 0: \inf_{\phi} \sup_{f_0 \in \mc{F}_{\epsilon}} \Ebb_{f_0}[1 - \phi] \leq \alpha \biggr\},
 \end{equation*} 
-where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation under the regression function $f_0$.\footnote{ Clearly, the minimax critical radius $\epsilon$ depends on $\alpha$.
-However we adopt the typical convention of treating $\alpha$ as a small but fixed positive constant; hence it will not affect the testing error rates, and we suppress it notationally.} See \citet{ingster82, ingster87, ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
+where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Of course, the minimax critical radius $\epsilon$ depends on $\alpha$.
+However, we adopt the typical convention of treating $\alpha$ as a small but fixed positive constant; hence it will not affect the testing error rates, and we suppress it notationally.} See \citet{ingster82,ingster87,ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
 
-Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing critical radius over $H^1(\Xset,M)$ is much smaller than the minimax estimation rate \citep{ingster2009}:
+Testing $f_0=0$ is an easier problem than estimating $f_0$, and hence the minimax testing critical radius over $H^1(\Xset,M)$ is smaller than the minimax estimation rate, for $1 \leq d < 4$ (see \citet{ingster2009}):
 \begin{equation}
 \label{eqn:sobolev_space_testing_critical_radius}
-\epsilon^2\Bigl(H^1(\Xset,M)\Bigr) \asymp M^{2d/(4 + d)}n^{-4/(4 + d)}~~\textrm{for $1 \leq d < 4$.}
+\epsilon^2\bigl(H^1(\Xset,M)\bigr) \asymp M^{2d/(4 + d)}n^{-4/(4 + d)}.
 \end{equation}
-When $d \geq 4$ the functions in $H^1(\Xset)$ are very irregular---formally speaking $H^1(\Xset)$ does not continuously embed into $\Leb^4(\Xset)$ when $d \geq 4$---and the minimax testing rates in this regime are unknown.
+When $d \geq 4$ the functions in $H^1(\Xset)$ are very irregular; formally speaking $H^1(\Xset)$ does not continuously embed into $\Leb^4(\Xset)$ when $d \geq 4$, and the minimax testing rates in this regime are unknown.
 
@@
 
-We now formalize the main conclusions of this work: that Laplacian smoothing methods on neighborhood graphs are minimax rate-optimal over first-order continuum Sobolev classes.
-Throughout this section, we use $C,C_1,\ldots$ and $c,c_1,\ldots$ to refer to constants which do not depend on $n$ or $f_0$.
-We will assume~\ref{asmp:domain} and ~\ref{asmp:bounded_lipschitz_density}, and additionally that the kernel $K$ satisfies condition~\ref{asmp:kernel}.
+We now formalize the main conclusions of this paper: that Laplacian smoothing methods on neighborhood graphs are minimax rate-optimal over first-order continuum Sobolev classes.
+We will assume \ref{asmp:domain}, \ref{asmp:density} on $P$, and the following condition on the kernel $K$.
 \begin{enumerate}[label=(K\arabic*)]
@@
 	\label{asmp:kernel}
-	$K:[0,\infty) \to [0,\infty)$ is a non-increasing function supported on $[0,1]$, its restriction to $[0,1]$ is Lipschitz, and $K(1) > 0$. Additionally, it is normalized so that
+	$K:[0,\infty) \to [0,\infty)$ is a nonincreasing function supported on $[0,1]$, its restriction to $[0,1]$ is Lipschitz, and $K(1) > 0$. Additionally, it is normalized so that
 	\begin{equation*}
-	\int_{\Reals^d} K\bigl(\norm{z}\bigr) \,dz = 1.
+	\int_{\Rd} K(\|z\|_2) \,dz = 1.
 	\end{equation*}
-	Let $\sigma_K := \frac{1}{d} \int_{\Rd} \|x\|^2 K(\|x\|) \,dx$.
+	We assume \smash{$\sigma_K = \frac{1}{d} \int_{\Rd} \|x\|_2^2 K(\|x\|_2) \,dx < \infty$}.
 \end{enumerate}
-This is a relatively mild condition: the choice of kernel is under the user's control, and moreover~\ref{asmp:kernel} covers many (though not all) common kernel choices.
+This is a mild condition: recall the choice of kernel is under the control of the user, and moreover \ref{asmp:kernel} covers many common kernel choices.
 
-\paragraph{Estimation error of Laplacian smoothing.} 
-Under these conditions, the Laplacian smoothing estimator $\wh{f}$ achieves, up to constants, the minimax risk $n^{-2/(2 + d)}$ over $H^1(\Xset,M)$.
+\paragraph{Estimation Error of Laplacian Smoothing.}
+
+Under these conditions, the Laplacian smoothing estimator \smash{$\wh{f}$} achieves an error rate that matches the minimax lower bound over $H^1(\Xset,M)$.
 This statement will hold whenever the graph $G_{n,r}$ is computed with radius $r$ in the following range.
@@
 	\label{asmp:ls_kernel_radius_estimation}
-	The neighborhood graph radius $r$ satisfies
+	For constants $C_0,c_0>0$, the neighborhood graph radius $r$ satisfies
 	\begin{equation*}
-	C_0\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq r \leq n^{-3/(4 + 2d)} M^{(d - 4)/(4 + 2d)} \wedge c_0.
+	C_0 \biggl(\frac{\log n}{n}\biggr)^{\frac{1}{d}} \leq r \leq c_0 \wedge M^{\frac{d - 4}{4 + 2d}} n^{-\frac{3}{4 + 2d}}.  
 	\end{equation*}
 \end{enumerate}
-We comment on~\ref{asmp:ls_kernel_radius_estimation} after stating Theorem~\ref{thm:laplacian_smoothing_estimation1}.
-The proof of this theorem, as with the proofs of all results in this paper, can be found in the appendices.
+
+Next we state Theorem~\ref{thm:laplacian_smoothing_estimation1}, our main estimation result. Its proof, as with all proofs of results in this paper, can be found in the appendix. 
 \begin{theorem}
 	\label{thm:laplacian_smoothing_estimation1}
-	Let $f_0 \in H^1(\Xset,M)$ with $d < 4$.
-Suppose that that the neighborhood graph $G_{n,r}$ is computed with a radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_estimation},  and the Laplacian smoothing estimator $\wh{f}$ with $\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$.
-With probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c \min\{M^{d/(2d + 4)} n^{d/(2+d)},n\})$, it holds that
-	\begin{equation}
-	\label{eqn:laplacian_smoothing_estimation1}
-	\Bigl\|\wh{f} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(2 + d)} n^{-2/(2 + d)}.
-	\end{equation}
+	Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ has dimension $d < 4$ and $M \leq n^{1/d}$.
+Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume the neighborhood graph \smash{$G_{n,r}$} is computed with a kernel $K$ satifsying \ref{asmp:kernel}.
+There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_estimation}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$}, satisfies
+	\begin{equation*}
+	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(2 + d)} n^{-2/(2 + d)},
+	\end{equation*}
+	with probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c (M^2n)^{d/(2 + d)})$.
+	% RJT: Alden, please verify that what I wrote is correct (I reorganized the way we explained this, added for $n \geq N$ ...).
+	% AJG: Verified.
 \end{theorem}
-To summarize: when $d = 1,2$ or $3$, with high probability the Laplacian smoothing estimator $\wh{f}$ has in-sample mean squared error within a constant factor of the minimax risk. Some remarks:
+
+To summarize: for $d = 1,2,$ or $3$, with high probability, the Laplacian smoothing estimator \smash{$\wh{f}$} has in-sample mean squared error that is within a constant factor of the minimax error.
+Some remarks:
 \begin{itemize}
-	\item The lower bound on $r$ imposed by~\ref{asmp:ls_kernel_radius_estimation} is  compatible with practice, where by far the most common choice of radius is the connectivity threshold $r \asymp (\log(n)/n)^{1/d}$, chosen to make the graph $G_{n,r}$ as sparse as possible while still being connected, for maximum computational efficiency.
-The upper bound is a bit more mysterious.
-We need it for technical reasons to ensure that $\wh{f}$ does not overfit, but note that as a practical matter one rarely chooses $r$ to be so large.
-	\item It is possible to extend $\wh{f}$ to be defined on all of $\Xset$, and evaluate the error of such an extension in $\Leb^2(\Xset)$ norm.
-Assuming that the extension of $\wh{f}$ and the regression function $f_0$ are both suitably smooth, tools from empirical process theory (Chapter 14 of \citet{wainwright2019}) or approximation theory (Section 15.5 of \citet{johnstone2011}) should guarantee that the $\Leb^2(\Xset)$ error is not too much greater than its in-sample counterpart.
-In fact, as we show in Section~\ref{subsec:laplacian_smoothing_estimation1_pf}, if $f_0$ is Lipschitz smooth and we extend $\wh{f}$ to be piecewise constant over the Voronoi tesselation induced by $X_1,\ldots,X_n$, the out-of-sample error $\|f_0 - \wh{f}\|_{\Leb^2(X)}$ is within a negligible factor of the in-sample error $\|f_0 - \wh{f}\|_n$, but we leave analysis of the Sobolev case to future work.
-	\item Under the stronger assumption that $f_0$ is Lipschitz smooth, we can replace the factor of $\delta$ in the high probability bound by the sharper $\delta^2/n$.
+	\item The first-order Sobolev space $H^1(\Xset)$ does not continuously embed into $C^0(\Xset)$ when $d>1$ (in general, the $k$th order space $H^k(\Xset)$ does not continuously embed into $C^0(\Xset)$ except if $2k>d$).
+For this reason, one really cannot speak of pointwise evaluation of a Sobolev function $f_0 \in H^1(\Xset)$ when $d>1$ (as we do in Theorem~\ref{thm:laplacian_smoothing_estimation1} by defining our target of estimation to be $f_0(X_i)$, $i=1,\ldots,n$).
+We can resolve this by appealing to what are known as \emph{Lebesgue points}, as explained in the appendix.
+	% RJT: Alden, please add a section to the supplement to explain this part.
+	% AJG: Added.
+	\item The assumption $M \leq n^{1/d}$ ensures that the upper bound provided in the theorem is meaningful (i.e., ensures it is of most a constant order).
+	\item The lower bound on $r$ imposed in condition \ref{asmp:ls_kernel_radius_estimation} is compatible with practice, where by far the most common choice of radius is the connectivity threshold \smash{$r \asymp (\log(n)/n)^{1/d}$}, which makes \smash{$G_{n,r}$} as sparse as possible while still being connected, for maximum computational efficiency.
+The upper bound may seem a bit more mysterious---we need it for technical reasons to ensure that \smash{$\wh{f}$} does not overfit, but note that as a practical matter one rarely chooses $r$ to be so large anyway.
+	\item It is possible to extend \smash{$\wh{f}$} to be defined on all of $\Xset$ and then evaluate the error of such an extension (as measured against $f_0$) in $\Leb^2(\Xset)$ norm.
+When \smash{$\wh{f}$} and $f_0$ are suitably smooth, tools from empirical process theory (see e.g., Chapter 14 of \citet{wainwright2019}) or approximation theory (e.g., Section 15.5 of \citet{johnstone2011}) guarantee that the $\Leb^2(\Xset)$ error is not too much greater than its in-sample counterpart.
+In fact, as shown in Appendix~\ref{sec:graph_sobolev_seminorm}, if $f_0$ is Lipschitz smooth and we extend \smash{$\wh{f}$} to be piecewise constant over the Voronoi tesselation induced by $X_1,\ldots,X_n$, then the out-of-sample error \smash{$\|\wh{f}-f_0\|_{\Leb^2(\Xset)}$} is within a negligible factor of the in-sample error \smash{$\|\wh{f}-f_0\|_n$}.
+We leave analysis of the Sobolev case to future work.
+	\item When $f_0$ is Lipschitz smooth, we can also replace the factor of $\delta$ in the high probability bound by the sharper $\delta^2/n$.
 \end{itemize}
 
-When $d = 4$, by setting $r \asymp (\log n/n)^{1/4}$ our analysis results in an upper bound on the mean squared error within a $(\log n)^{1/3}$ factor of the minimax risk, but when $d \geq 5$ our analysis yields upper bounds that are much worse than the minimax rate.
-This mirrors the conclusions of~\citet{sadhanala16}, who investigate estimation rates of Laplacian smoothing over the $d$-dimensional grid graph.
-\citet{sadhanala16} argue that their analysis is tight, and that it is the estimator, rather than the analysis, which is deficient when $d \geq 5$.
-Formally proving such a claim turns out to be harder in the random design setting than in the fixed design setting, and we leave it for future work.
+When $d = 4$, our analysis results in an upper bound for the error of Laplacian smoothing that is within a \smash{$(\log n)^{1/3}$} factor of the minimax error rate.
+But when $d \geq 5$, our upper bounds are much worse.
+\begin{theorem}
+	\label{thm:laplacian_smoothing_estimation2}
+	Under the assumptions of Theorem~\ref{thm:laplacian_smoothing_estimation1}, if instead $\Xset$ has dimension $d = 4$, \smash{$r \asymp (\log n/n)^{1/4}$} and \smash{$\rho = M^{-2/3}(nr^6)^{-1}(\log n/n)^{1/3}$}, then we obtain
+	\begin{equation*}
+	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{4/3} \biggl(\frac{\log n}{n}\biggr)^{1/3},
+	\end{equation*}
+	with the same probability guarantee as in Theorem~\ref{thm:laplacian_smoothing_estimation1}.
+If the dimension of $\Xset$ is $d \geq 5$, \smash{$r \asymp (\log n/n)^{1/d}$} and \smash{$\rho =  M^{-2/3}(nr^{2 + d})^{-1}n^{-4/(3d)}$}, then
+	\begin{equation*}
+	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{4/3} \biggl(\frac{\log n}{n}\biggr)^{-4/(3d)},
+	\end{equation*}
+	again with the same probability guarantee.
+	% RJT: Alden, please verify what I wrote is correct, and fill in the result for d \geq 5.
+	% AJG: Verified and filled in.
+\end{theorem}
+This mirrors the conclusions of \citet{sadhanala16} who investigate estimation rates of Laplacian smoothing over the $d$-dimensional grid graph.
+These authors argue that their analysis is tight, and that it is likely the estimator, not the analysis, that is deficient when $d \geq 5$.
+Formalizing such a claim turns out to be harder in the random design setting than in the fixed design setting, and we leave it for future work.
 
 However, we do investigate the matter empirically.
-In Figure~\ref{fig:fig1}, we consider the mean squared error of the Laplacian smoothing estimator, for several different values of dimension $d$.
-Here, $X_1,\ldots,X_n$ are sampled uniformly over $\Xset = [-1,1]^d$, and the regression function $f_0(x) \propto \Pi_{i = 1}^{d} \cos(m \pi x_i)$, where $m = 2$ for $d = 2$, and $m = 1$ for $d \geq 3$.
-This is a rather smooth regression function, and when $d = 2$ and $d = 3$ empirically Laplacian smoothing appears to achieve or exceed the minimax rate.
-When $d = 4$ Laplacian smoothing appears modestly suboptimal; this fits with our theoretical upper bound, which includes a $(\log n)^{1/3}$ factor that plays a non-negligible role for these problem sizes.
-On the other hand, when $d = 5$ Laplacian smoothing seems to be decidedly suboptimal.
+In Figure~\ref{fig:fig1}, we study the (in-sample) mean squared error of the Laplacian smoothing estimator as the dimension $d$ grows.
+Here $X_1,\ldots,X_n$ are sampled uniformly over $\Xset = [-1,1]^d$, and the regression function is taken as \smash{$f_0(x) \propto \Pi_{i = 1}^{d} \cos(a \pi x_i)$}, where $a = 2$ for $d = 2$, and $a = 1$ for $d \geq 3$.
+This regression function $f_0$ is quite smooth, and for $d = 2$ and $d = 3$ Laplacian smoothing appears to achieve or exceed the minimax rate.
+When $d = 4$, Laplacian smoothing appears modestly suboptimal; this fits with our theoretical upper bound, which includes a \smash{$(\log n)^{1/3}$} factor that plays a non-negligible role for these problem sizes ($n=1000$ to $n=10000$).
+On the other hand, when $d = 5$, Laplacian smoothing seems to be decidedly suboptimal.
 
-\begin{figure}[thb]
-	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_2d.pdf}
-	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_3d.pdf}
-	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_4d.pdf}
-	\includegraphics[width=.23\textwidth]{figures/cosine/mse_by_sample_size_5d.pdf}
+\begin{figure*}[tb]
+	\includegraphics[width=.245\textwidth]{figures/cosine/mse_by_sample_size_2d.pdf}
+	\includegraphics[width=.245\textwidth]{figures/cosine/mse_by_sample_size_3d.pdf} 
+	\includegraphics[width=.245\textwidth]{figures/cosine/mse_by_sample_size_4d.pdf}
+	\includegraphics[width=.245\textwidth]{figures/cosine/mse_by_sample_size_5d.pdf}
+	\caption{Mean squared error of Laplacian smoothing (\texttt{LS}) as a function of sample size $n$.
+Each plot is on the log-log scale, and the results are averaged over 5 repetitions, with Laplacian smoothing tuned for optimal average mean squared error.
+The black line shows the minimax rate (in slope only; the intercept is chosen to match the observed error).}
+	\label{fig:fig1}
+	% RJT: Alden, please verify that what I wrote about the intercept above is right.
+And if so, can you redo the black line for d = 2 so that the intercept matches better?
+Thanks.
+Oh and if you're already going to redo one of these plots, you can satisfy my unreasonably nitpicky brain and change the x- and y-axes on all of them to use words rather than symbols/acronyms: "Sample size" and "Mean squared error".
+	% AJG: Verified, and redone with requested changes.
+\end{figure*}
 
-	\caption{Mean squared error of Laplacian smoothing (\texttt{LS}) as a function of sample size $n$.
-Plots are on the log-log scale, MSE is averaged over 5 iterations, with Laplacian smoothing tuned for optimal MSE.
-The black line shows the minimax rate.}
-	\label{fig:fig1}
-\end{figure}
+\paragraph{Testing Error of Laplacian Smoothing.}
 
-\paragraph{Testing error of Laplacian smoothing.}
-Let us define a test using the statistic $\wh{T}$. For $b \geq 1$, set the threshold $\wh{t}_b$ as
+For $b \geq 1$, define a threshold \smash{$\wh{t}_b$} as
 \begin{equation*}
-\wh{t}_{b} := \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k + 1\bigr)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k + 1\bigr)^4}},
+\wh{t}_{b} = \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}},
 \end{equation*}
-where we recall $\lambda_k$ is the $k$th smallest eigenvalue of $\Lap_{n,r}$; the Laplacian smoothing test is simply $\wh{\varphi} := \1\bigl\{\wh{T} \leq \wh{t}_b\bigr\}$.
-(Clearly the test $\wh{\varphi}$ depends on $b$, but we suppress this notationally.) The number $b \geq 1$ is a tuning parameter selected by the user, with the choice made based on the tolerated level of Type I and Type II error.
-In fact, the threshold $\wh{t}_b$ is precisely the right choice to control the Type I error of $\wh{\varphi}$; as we show in Section~\ref{sec:fixed_graph_error_bounds},
+where we recall $\lambda_k$ is the $k$th smallest eigenvalue of \smash{$\Lap_{n,r}$}. The Laplacian smoothing test is then simply
+\begin{equation*}
+\wh{\varphi} = \1\bigl\{\wh{T} \leq \wh{t}_b\bigr\}.
+\end{equation*} 
+%(Clearly \smash{$\wh{\varphi}$} depends on $b$, but we suppress this notationally.) 
+Here $b \geq 1$ is a tuning parameter selected by the user, with the choice driven by the tolerated Type I versus Type II error.
+In fact, the threshold \smash{$\wh{t}_b$} is precisely the right choice to control the Type I error of \smash{$\wh{\varphi}$}: we show in Appendix~\ref{sec:fixed_graph_error_bounds} that
 \begin{equation}
@@
 \end{equation}
-Theorem~\ref{thm:laplacian_smoothing_testing} establishes that this test also has small Type II error, uniformly over all $f_0$ separated from $0$ by at least the critical radius $\epsilon\bigl(H^1(\Xset,M)\bigr)$ given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
-For this to hold, we will require a tight range of scalings for $r$.
+The next theorem shows that this test also has small Type II error, uniformly over all $f_0$ that are separated from $0$ by at least the critical radius given in \eqref{eqn:sobolev_space_testing_critical_radius}.
+For this to hold, we will require a tighter range of scalings for the graph radius $r$.
 \begin{enumerate}[label=(R\arabic*)]
@@
 	\label{asmp:ls_kernel_radius_testing}
-	The neighborhood graph radius $r$ satisfies
+	For constants $C_0,c_0>0$, the neighborhood graph radius $r$ satisfies
 	\begin{equation*}
-	C_0\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq r \leq M^{\frac{(d - 8)}{8 + 2d}}n^{\frac{d - 20}{32 + 8d}} \wedge c_0.
+	C_0\biggl(\frac{\log n}{n}\biggr)^{\frac{1}{d}} \leq r \leq c_0 \wedge M^{\frac{(d - 8)}{8 + 2d}} n^{\frac{d - 20}{32 + 8d}}.
 	\end{equation*}
 \end{enumerate}
-Note that when $d < 4$, if $M$ is not too large there always exists a choice of $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}. Put
+We will also require that the radius of the graph Sobolev class not be too large. Precisely, we will require $M \leq M_{\max}(d)$, where we define
 \begin{equation*}
@@
 \begin{cases*}
-n^{1/8},& \textrm{$d = 1$}\\
-n^{(4 - d)/(4d)},& \textrm{$d \geq 2$}.
+n^{1/8} & \textrm{$d = 1$} \\
+n^{(4 - d)/(4d)} & \textrm{$d \geq 2$}.
 \end{cases*}
 \end{equation*}
+
+We now give Theorem~\ref{thm:laplacian_smoothing_testing}, our main testing result.
 \begin{theorem}
 	\label{thm:laplacian_smoothing_testing}
-	Let $f_0 \in H^1(\Xset,M)$ for $M \leq M_{\max}(d)$ and $d < 4$.
-Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing}, and the Laplacian smoothing test $\wh{\varphi}$ with $\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$, and $b \geq 1$.
-For all $f_0$ such that
+	Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ with $d < 4$, and where $M \leq M_{\max}(d)$.
+Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume \smash{$G_{n,r}$} is computed with a kernel $K$ satisfying \ref{asmp:kernel}.
+There are constants $N,C,c>0$ such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_testing}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$}, satisfies the following: for all $b \geq 1$, and all $f_0$ such that
 	\begin{equation}
 	\label{eqn:laplacian_smoothing_testing}
-	\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2d/(4 + d)} n^{-4/(4 + d)}
+	\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2d/(4 + d)} n^{-4/(4 + d)},
 	\end{equation} 
-	the Type II error is upper bounded:
-	\begin{align*}
-	& \Ebb_{f_0}\Bigl[1 - \wh{\varphi} \Bigr] \leq \\
-	& C\biggl(\frac{1}{b}\Bigl[1 + n^{-d/(4 + d)}M^{-2d/(4 + d)}\Bigr] + n\exp\bigl(-c_1nr^d\bigr)\biggr).
-	\end{align*}
+	the Type II error is upper bounded by
+	\begin{equation*}
+	\Ebb_{f_0}\bigl[1 - \wh{\varphi} \bigr] \leq C\biggl(\frac{1}{b}\Bigl[1 + n^{-d/(4 + d)}M^{-2d/(4 + d)}\Bigr] + n\exp\bigl(-cnr^d\bigr)\biggr).
+	\end{equation*}
 \end{theorem}
-A couple remarks:
+
+Some remarks:
 \begin{itemize}
-	\item As mentioned in Section~\ref{sec:problem_setup_and_background}, when $d \geq 4$ the Sobolev balls $H^1(\Xset,M)$ include quite irregular functions $f \not\in \Leb^4(\Xset)$.
-Proving tight lower bounds over such classes is non-trivial, and to the best of our knowledge such an analysis remains outstanding.
-On the other hand, if we explicitly assume that $f_0 \in \Leb^4(\Xset,M)$, then \citet{guerre02} show that the testing problem is characterized by the dimension-free lower bound $\epsilon^{2}(\Leb^4(\Xset,M)) \gtrsim n^{-1/2}$.
-Moreover, by training $\wh{f}$ to the interpolation limit, i.e.
-setting $\rho = 0$, the test $\wh{\varphi}$ achieves (up to constants) this lower bound.
-That is, for any $f_0 \in \Leb^4(\Xset,M)$ such that $\|f_0\|_{\Leb^2(\Xset)}^2 \geq C b^2n^{-1/2}$, we have both
+	\item As mentioned earlier, Sobolev balls $H^1(\Xset,M)$ for $d \geq 4$ include quite irregular functions $f \not\in \Leb^4(\Xset)$.
+Proving tight lower bounds in this case is nontrivial, and as far as we understand such an analysis remains outstanding.
+On the other hand, if we explicitly assume that $f_0 \in \Leb^4(\Xset,M)$, then \citet{guerre02} show that the testing problem is characterized by a dimension-free lower bound $\epsilon^{2}(\Leb^4(\Xset,M)) \gtrsim n^{-1/2}$.
+Moreover, by training \smash{$\wh{f}$} to the interpolation limit, setting $\rho = 0$, the subsequent test \smash{$\wh{\varphi}$} will achieve (up to constants) this lower bound.
+That is, for any $f_0 \in \Leb^4(\Xset,M)$ such that \smash{$\|f_0\|_{\Leb^2(\Xset)}^2 \geq C b^2n^{-1/2}$}, we have both
 	\begin{equation}
 	\label{eqn:laplacian_smoothing_testing_low_smoothness}
-	E_{f_0}\Bigl[1 - \wh{\varphi}\Bigr] \leq \frac{C(1 + M^4)}{b^2}
+	\Ebb_{f_0}\bigl[1 - \wh{\varphi}\bigr] \leq \frac{C(1 + M^4)}{b^2},
 	\end{equation} 
-	and $\Ebb_0[\wh{\varphi}] \leq 1/b^2$. These same results apply to $\wt{T}$, since when $d > 1$ the estimator $\wt{f}$ will interpolate the responses $Y_1,\ldots,Y_n$.
-	\item To compute the data-dependent threshold $\wh{t}_b$, one must know each of the eigenvalues $\lambda_1,\ldots,\lambda_n$.
-Computing all $n$ of these eigenvalues is far more expensive---on the order of $n^3$ operations---than computing $\wh{T}$.
-That being said, in practice we would not recommend using $\wh{t}_b$ anyway, and would instead we make the standard recommendation to calibrate via a permutation test \citep{hoeffding1952}.
+	and \smash{$\Ebb_0[\wh{\varphi}] \leq 1/b^2$}.
+Note that the same results apply to \smash{$\wt{T}$}, since the thin-plate spline estimator \smash{$\wt{f}$} interpolates the responses $Y_1,\ldots,Y_n$ for $d>1$.
+	\item To compute the data-dependent threshold \smash{$\wh{t}_b$}, one must know all of the eigenvalues $\lambda_1,\ldots,\lambda_n$.
+Computing all these eigenvalues is far more expensive (cubic-time) than computing \smash{$\wh{T}$} in the first place (nearly-linear-time).
+But in practice we would not recommend using \smash{$\wh{t}_b$} anyway, and would instead we make the standard recommendation to calibrate via a permutation test \citep{hoeffding1952}.
 \end{itemize}
 
-\paragraph{Comparison with thin-plate splines.}
-With some of our main results in hand, let us pause to offer some explanation of why Laplacian smoothing can be optimal in settings where thin-plate splines are not even consistent.
+\paragraph{More Discussion of Variational Analog.}
 
-First, we should make clear why this difference in performance is so surprising.
-As mentioned previously, the penalties in~\eqref{eqn:laplacian_smoothing} and~\eqref{eqn:thin_plate_spline} can be closely tied together: \citet{bousquet03} show that for any $f \in C^2(\Xset)$,
+With some results in hand, let us pause to offer some explanation of why Laplacian smoothing can be optimal in settings where thin-plate splines are not even consistent.
+First, we elaborate on why this difference in performance is so surprising.
+As mentioned previously, the penalties in \eqref{eqn:laplacian_smoothing}, \eqref{eqn:thin_plate_spline} can be closely tied together: \citet{bousquet03} show that for $f \in C^2(\Xset)$,
 \begin{equation}
@@
 \begin{aligned}
-\lim \frac{1}{n^2 r^{d + 2}} f^T \Lap_{n,r} f & = \int_{\Xset} f(x) \cdot \Delta_Pf(x) p(x) \,dx \\
-& = \int_{\Xset} \|\nabla f(x)\|^2 p^2(x) \,dx.
+\lim \frac{1}{n^2 r^{d + 2}} f^\top \Lap_{n,r} f & = \int_{\Xset} f(x) \cdot \Delta_Pf(x) p(x) \,dx \\
+& = \int_{\Xset} \|\nabla f(x)\|_2^2 p^2(x) \,dx.
 \end{aligned}
@@
 \begin{equation*}
-\Delta_Pf := -\frac{1}{p} \dive\bigl(p^2\nabla f)
+\Delta_Pf := -\frac{1}{p} \dive\bigl(p^2\nabla f),
 \end{equation*}
-and the second equality follows from integration by parts.\footnote{Assuming $f$ satisfies e.g.
-Dirichlet boundary conditions.} To be clear, this argument does not formally imply that $\wh{f}$ and $\wt{f}$ are close---note that~\eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:thin_plate_spline} is $H^1(\Xset)$---but it does seem to suggest that the two estimators should behave somewhat similarly.
+and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies e.g., Dirichlet boundary conditions.} To be clear, this argument does not formally imply that $\wh{f}$ and $\wt{f}$ are close (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of continuous functions with weak derivatives in $\Leb^2(\Xset)$).
+But it does seem to suggest that the two estimators should behave somewhat similarly.
 
-Of course, we know this is not the case: $\wh{f}$ and $\wt{f}$ look very different when $d > 1$.
+Of course, we know this is not the case: \smash{$\wh{f}$} and \smash{$\wt{f}$} look very different when $d > 1$.
 What is driving this difference?
-The key point is that the discretization imposed by the graph $G_{n,r}$---which seemed so problematic at first glance---turns out to be a blessing.
-The problem with~\eqref{eqn:thin_plate_spline} is that the function class $H^1(\Xset)$ is far too large when $d \geq 2$.
+The key point is that the discretization imposed by the graph $G_{n,r}$---which might seem problematic at first glance---turns out to be a blessing.
+The problem with~\eqref{eqn:thin_plate_spline} is that the class $H^1(\Xset)$, which fundamentally underlies the criterion, is far ``too big'' for $d > 1$.
 This is meant in various related senses.
-From the Sobolev embedding theorem \citep{evans10}, we know that $H^1(\Xset)$ does not continuously embed into any H\"{o}lder space, and that $H^1(\Xset)$ is not a Reproducing Kernel Hilbert Space.
-More relevant to this context, by placing ``bump functions'' of arbitrarily small radius around each of the design points $X_1,\ldots,X_n$, one can interpolate the responses $Y_1,\ldots,Y_n$ in such a manner that forces the criterion in~\eqref{eqn:thin_plate_spline} to 0 \citep{green93}; the resulting estimator will clearly be inconsistent.
+By the Sobolev embedding theorem, for $d>1$, the class $H^1(\Xset)$ does not continuously embed into any H\"{o}lder space; and in fact it does not even continuously embed into $C^0(\Xset)$.
+Thus we cannot really restrict the optimization to \emph{continuous} and weakly differentiable functions, as we could when $d=1$ (the smoothing spline case), without throwing out a substantial subset of functions in $H^1(\Xset)$.
+Even among continuous and differentiable functions $f$, as we explained previously, we can use ``bump'' functions (as in \citet{green93}) to construct $f$ that interpolates the pairs $(X_i,Y_i)$, $i=1,\ldots,n$ and achieves arbitrarily small penalty (and hence criterion) in \eqref{eqn:thin_plate_spline}.
+In this sense, any estimator resulting from solving \eqref{eqn:thin_plate_spline} will clearly be inconsistent.
 
-On the other hand~\eqref{eqn:laplacian_smoothing} is a finite dimensional (if growing linearly with $n$) problem---as a result $\wh{f}$ has far less capacity to overfit than does $\wt{f}$.
-Of course, discretization is not the only way to make the problem~\eqref{eqn:thin_plate_spline} more tractable: for instance, one could replace the penalty $\int_{\Xset} \|\nabla f(x)\|^2 \,dx$ by the stricter choice $\sup_{x \in \Xset} \|\nabla f(x)\|$,  or conduct the problem over some finite dimensional linear subspace of $H^1(\Xset)$ (i.e use a sieve).
-These solutions improve the statistical properties of $\wt{f}$ when $d > 1$ \citep{birge1993,birge1998,vandergeer2000}, but the first is computationally intractable, while the second requires the estimator to be specifically tailored to the domain $\Xset$, in stark contrast to $\wh{f}$.
+On the other hand, problem \eqref{eqn:laplacian_smoothing} is finite-dimensional.
+As a result \smash{$\wh{f}$} has far less capacity to overfit than does \smash{$\wt{f}$}, for any given sample size $n$.
+Discretization is not the only way to make the problem \eqref{eqn:thin_plate_spline} more tractable: for instance, one can replace the penalty \smash{$\int_{\Xset} \|\nabla f(x)\|_2^2 \,dx$} with a stricter choice like \smash{$\esssup_{x \in \Xset} \|\nabla f(x)\|_2$}, or conduct the optimization over some finite-dimensional linear subspace of $H^1(\Xset)$ (i.e., use a sieve).
+While these solutions do improve the statistical properties of \smash{$\wt{f}$} for $d > 1$ (see e.g., \citet{birge1993,birge1998,vandergeer2000}), Laplacian smoothing is generally speaking much simpler and more computationally friendly.
+In addition, the other approaches are specifically tailored to the domain $\Xset$, in stark contrast to \smash{$\wh{f}$}.
 
-\paragraph{Overview of analysis.}
-The comparison with thin-plate splines highlights some surprising strengths of $\wh{f}$.
-Unfortunately, this comparison also precludes analyzing $\wh{f}$ by using~\eqref{eqn:seminorm_consistency} to establish a coupling between $\wh{f}$ and $\wt{f}$; we know this won't work, because we would like to prove tight error bounds on $\wh{f}$ in regimes where no such bounds exist for $\wt{f}$.
+\paragraph{Overview of Analysis.}
 
-Instead we take a different approach, and directly analyze the error of $\wh{f}$ and $\wh{T}$ using a (conditional on $X_1,\ldots,X_n$) bias-variance decomposition. A standard calculation shows that
+The comparison with thin-plate splines highlights some surprising differences between \smash{$\wh{f}$} and \smash{$\wt{f}$}.
+Such differences also preclude us from analyzing \smash{$\wh{f}$} by, say, using \eqref{eqn:seminorm_consistency} to establish a coupling between \smash{$\wh{f}$} and \smash{$\wt{f}$}---we know this cannot work, because we would like to prove meaningful error bounds on \smash{$\wh{f}$} in regimes where no such bounds exist for \smash{$\wt{f}$}.
+
+Instead we take a different approach, and directly analyze the error of \smash{$\wh{f}$} and \smash{$\wh{T}$} using a bias-variance decomposition (conditional on $X_1,\ldots,X_n$).
+A standard calculation shows that
 \begin{equation*}
-\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \underbrace{\frac{2\rho}{n} \bigl(f_0^T \Lap_{n,r} f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{10}{n} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2}}_{\textrm{variance}}
+\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \underbrace{\vphantom{\sum_{k=1}^n}\frac{2\rho}{n} \bigl(f_0^\top \Lap_{n,r} f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{10}{n} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2}}_{\textrm{variance}},
 \end{equation*}
@@
 \begin{equation*}
-\bigl\|f_0\bigr\|_n^2 \geq \underbrace{\frac{2\rho}{n} \bigl(f_0^T \Lap_{n,r} f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{4b}{n} \sqrt{\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}}}_{\textrm{variance}}
+\bigl\|f_0\bigr\|_n^2 \geq \underbrace{\vphantom{\sqrt{\sum_{k=1}^n}}\frac{2\rho}{n} \bigl(f_0^\top \Lap_{n,r} f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{4b}{n} \sqrt{\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}}}_{\textrm{variance}}.
 \end{equation*}
-Note that the bias and variance terms are each functionals of the random graph $G_{n,r}$, and so are themselves random.
-To upper bound them, we build on some recent works \citep{burago2014,trillos2019,calder2019} regarding the consistency of neighborhood graphs to establish the following two Lemmas.
-In the second of these Lemmas, we will assume
-\begin{enumerate}[label=(R\arabic*)]
-	\setcounter{enumi}{2}
-	\item
-	\label{asmp:radius1}
-	The radius $r$ satisfies $C_0(\log n/n)^{1/d} \leq r \leq c_0$.
-\end{enumerate}
-and put $A_{n,r}(k) := \min\{nr^{d + 2}k^{2/d},nr^d\}$.
+The bias and variance terms are each functions of the random graph \smash{$G_{n,r}$}, and hence are themselves random.
+To upper bound them, we build on some recent works \citep{burago2014,trillos2019,calder2019} regarding the consistency of neighborhood graphs to establish the following lemmas.
+These lemmas assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and \ref{asmp:kernel} on the kernel used to compute the neighborhood graph \smash{$G_{n,r}$}.
 
@@
 	\label{lem:graph_sobolev_seminorm}
-	For any $f \in H^1(\Xset)$, with probability at least $1 - \delta$, it holds that
+	There are constants $N,C_2 > 0$ such that for $n \geq N$, $r \leq c_0$, and $f \in H^1(\Xset)$, with probability at least $1 - \delta$, it holds that
 	\begin{equation}
 	\label{eqn:graph_sobolev_seminorm}
-	f^T \Lap_{n,r} f \leq \frac{C_2}{\delta} n^2 r^{d + 2} |f|_{H^1(X)}^2.
-	\end{equation}
-\end{lemma}
-\begin{lemma}
-	\label{lem:neighborhood_eigenvalue}
-	For any neighborhood graph radius $r$ which satisfies~\ref{asmp:radius1}, with probability at least $1 - C_1n\exp(-c_1nr^d)$ it holds that
-	\begin{equation}
-	\label{eqn:neighborhood_eigenvalue}
-	c_3A_{n,r}(k) \leq \lambda_k \leq C_3A_{n,r}(k)~~\textrm{for all $2 \leq k \leq n$}
+	f^\top \Lap_{n,r} f \leq \frac{C_2}{\delta} n^2 r^{d + 2} |f|_{H^1(\Xset)}^2.
 	\end{equation}
@@
 
+\begin{lemma}
+	\label{lem:neighborhood_eigenvalue} 
+	There are constants $N,C_1,C_3,c_1,c_3 > 0$ such that for $n \geq N$ and $C_0(\log n/n)^{1/d} \leq r \leq c_0$, with probability at least $1 - C_1n\exp(-c_1nr^d)$, it holds that
+	\begin{equation}
+	\label{eqn:neighborhood_eigenvalue}
+	c_3A_{n,r}(k) \leq \lambda_k \leq C_3A_{n,r}(k), ~~\textrm{for $2 \leq k \leq n$},
+	\end{equation}
+	where \smash{$A_{n,r}(k) = \min\{nr^{d + 2}k^{2/d},nr^d\}$}.
+\end{lemma}
+
 Lemma~\ref{lem:graph_sobolev_seminorm} gives a direct upper bound on the bias term.
-Lemma~\ref{lem:neighborhood_eigenvalue} results in a sufficient upper bound on the variance term whenever the neighborhood graph radius $r$ is \textit{sufficiently small}; precisely, when $r$ is upper bounded as in~\ref{asmp:ls_kernel_radius_estimation} (for estimation) or~\ref{asmp:ls_kernel_radius_testing} (for testing).
+Lemma~\ref{lem:neighborhood_eigenvalue} leads to a sufficiently tight upper bound on the variance term whenever the radius $r$ is sufficiently small; precisely, when $r$ is upper bounded as in \ref{asmp:ls_kernel_radius_estimation} for estimation, or \ref{asmp:ls_kernel_radius_testing} for testing.
 The parameter $\rho$ is then tuned to minimize the sum of bias and variance, in the usual manner.
 
-It is useful to give another perspective on our approach.
-When analyzing penalized least squares estimators such as $\wh{f}$ or $\wt{f}$ one typically assumes two properties: first, that the regression function $f_0$ lies in (or near) a ball in the normed space induced by the penalty, and second that this ball is reasonably small, e.g.
-as measured by metric entropy.
-In contrast, the Laplacian smoothing penalty induces a ball
+It may be useful to give one more perspective on our approach.
+A common strategy in analyzing penalized least squares estimators is to assume two properties: first, that the regression function $f_0$ lies in (or near) a ball defined by the penalty operator; second, that this ball is reasonably small, e.g., as measured by metric entropy, or Rademacher complexity, etc.
+In contrast, in Laplacian smoothing, the penalty induces a ball
 \begin{equation*}
-H^1(G_{n,r},M) := \{f: f^T \Lap_{n,r} f \leq M^2\}
+H^1(G_{n,r},M) := \{f: f^\top \Lap_{n,r} f \leq M^2\}
 \end{equation*}
-that is data-dependent and random, and so we do not have access to either of the aforementioned properties \emph{a priori}; instead we must prove they hold with high probability.
-In this sense, our analysis is rather different than is typical in nonparametric regression.
+that is data-dependent and random, and so we do not have access to either of the aforementioned properties a priori, and instead, must prove they hold with high probability.
+In this sense, our analysis is different than the typical one in nonparametric regression.
 
@@
 \label{sec:manifold_adaptivity}
-The minimax rates $n^{-2/(2 + d)}$ (in estimation) and $n^{-4/(4 + d)}$ (in testing) suffer from the curse of dimensionality.
-However, in practice it is often reasonable to assume a \emph{manifold hypothesis}: roughly speaking, that the data $X_1,\ldots,X_n$ lie on a submanifold $\Xset$ of $\Rd$ that has intrinsic dimension $1 \leq m \leq d$.
-Under such an assumption, it is known \citep{bickel2007, ariascastro2018} that the optimal rates over $H^1(\Xset)$ now scale like $n^{-2/(2 + m)}$ (for estimation) and $n^{-4/(4 + m)}$ (for testing), which can be much smaller than the full-dimensional error rates when $m \ll d$.
 
-On the other hand, a theory has been developed~\citep{belkin03,belkin05,niyogi2013} establishing that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
-In this section, we contribute to this line of work by showing that under the manifold hypothesis, Laplacian smoothing achieves the sharper minimax rates over $H^1(\Xset)$.
+The minimax rates $n^{-2/(2 + d)}$ and $n^{-4/(4 + d)}$, in estimation and testing, suffer from the curse of dimensionality.
+However, in practice it can be often reasonable to assume a \emph{manifold hypothesis}: that the data $X_1,\ldots,X_n$ lie on a manifold $\Xset$ of $\Rd$ that has intrinsic dimension $m < d$.
+Under such an assumption, it is known \citep{bickel2007,ariascastro2018} that the optimal rates over $H^1(\Xset)$ are now $n^{-2/(2 + m)}$ (for estimation) and $n^{-4/(4 + m)}$ (for testing), which are much faster than the full-dimensional error rates when $m \ll d$.
 
-\paragraph{Error rates with the manifold hypothesis.}
-The conditions and results of this Section will be largely similar to the previous one, except with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
-For the rest of this section we assume the following.
+% RJT: Again, would venture a guess that these are stated results for Holder classes, not Sobolev classes.
+So same comment as before, as when we referenced Tsybakov.
+Just checking my understanding ...
+not sure it's worth changing this at this point.
+
+On the other hand, a theory has been developed \citep{belkin03,belkin05,niyogi2013} establishing that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+We contribute to this line of work by showing that under the manifold hypothesis, Laplacian smoothing achieves the sharper minimax rates over $H^1(\Xset)$.
+
+\paragraph{Error Rates Assuming the Manifold Hypothesis.}
+
+The conditions and results presented here will be largely similar to the previous ones, except with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
+For the remainder, we assume the following.
 \begin{enumerate}[label=(P\arabic*)]
@@
 	\label{asmp:domain_manifold}
-	The measure $P$ is supported on a compact, connected, smooth submanifold $\Xset$ of $\Reals^d$.
-The manifold $\Xset$ is of fixed dimension $1 \leq m \leq d$, is without boundary, and has positive reach \citep{federer1959}.
+	$P$ is supported on a compact, connected, smooth manifold $\Xset$ embedded in $\Rd$, of dimension $m \leq d$. The manifold is without boundary and has positive reach \citep{federer1959}.
 	\item 
-	\label{asmp:density_manifold} $P$ has a density $p$ with respect to the volume form of $\Xset$ which is bounded away from $0$ and $\infty$,
+	\label{asmp:density_manifold} 
+	$P$ admits a density $p$ with respect to the volume form of $\Xset$ such that 
 	\begin{equation*}
-	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty, ~~\textrm{for all $x \in \Xset$}. 
 	\end{equation*}
-	for all $x \in \Xset$. The density $p$ is Lipschitz with Lipschitz constant $L_p$.
+	Additionally, $p$ is Lipschitz on $\Xset$, with Lipschitz constant $L_p$.
 \end{enumerate}
-Under the assumptions~\ref{asmp:domain_manifold}, \ref{asmp:density_manifold} and~\ref{asmp:kernel}, and for a suitable range of $r$, the error bounds on the estimator $\wh{f}$ and test $\wh{\varphi}$ will depend on $m$ instead of $d$.
 
+Under the assumptions \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}, and \ref{asmp:kernel}, and for a suitable range of $r$, the error bounds on the estimator \smash{$\wh{f}$} and test \smash{$\wh{\varphi}$} will depend on $m$ instead of $d$.
 \begin{enumerate}[label=(R\arabic*)]
@@
 	\label{asmp:ls_kernel_radius_estimation_manifold}
-	The neighborhood graph radius $r$ satisfies
+	For constants $C_0,c_0>0$, the neighborhood graph radius $r$ satisfies 
 	\begin{equation*}
-	C_4\biggl(\frac{\log n}{n}\biggr)^{1/m} \leq r \leq n^{\frac{-3}{(4 + 2m)}} M^{\frac{(m - 4)}{(4 + 2m)}} \wedge c_4.
+	C_0\biggl(\frac{\log n}{n}\biggr)^{\frac{1}{m}} \leq r \leq c_0 \wedge  M^{\frac{(m - 4)}{(4 + 2m)}} n^{\frac{-3}{(4 + 2m)}}.
 	\end{equation*}
 \end{enumerate}
+
 \begin{theorem}
 	\label{thm:laplacian_smoothing_estimation_manifold}
-	Let $f_0 \in H^1(\Xset,M)$ with $m < 4$.
-Suppose that that the neighborhood graph $G_{n,r}$ is computed with a radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_estimation_manifold},  and the Laplacian smoothing estimator $\wh{f}$ with $\rho = M^{-4/(2 + m)} (nr^{m + 2})^{-1} n^{-2/(2 + m)}$.
-With probability at least $1 - \delta -  Cn\exp(-cnr^m) - \exp(-c M^{m/(2m + 4)} n^{m/(2+m)})$, it holds that
+	As in Theorem \ref{thm:laplacian_smoothing_estimation1}, but where $\Xset \subseteq \Rd$ is a manifold with intrinsic dimension $m < 4$, the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}, and $M \leq n^{1/m}$.
+There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any \smash{$n \geq N$}, and any $r$ as in \ref{asmp:ls_kernel_radius_estimation_manifold}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + m)} (nr^{m + 2})^{-1} n^{-2/(2 + m)}$}, satisfies
 	\begin{equation*}
-	\Bigl\|\wh{f} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2m/(2 + m)} n^{-2/(2 + m)}.
+	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{2m/(2 + m)} n^{-2/(2 + m)},
 	\end{equation*}
+	with probability at least $1 - \delta -  C_1 n\exp(-cnr^m) - \exp(-c(M^2n)^{m/(2+m)})$.
+	% RJT: Alden, please verify that what I wrote is correct.
+	% AJG: Verified.
 \end{theorem}
 
-We require the following range of $r$ for testing.
 \begin{enumerate}[label=(R\arabic*)]
@@
 	\label{asmp:ls_kernel_radius_testing_manifold}
-	The neighborhood graph radius $r$ satisfies
+	For constants $C_0,c_0>0$, the neighborhood graph radius $r$ satisfies 
 	\begin{equation*}
-	C_4\biggl(\frac{\log n}{n}\biggr)^{1/m} \leq r \leq M^{\frac{(m - 8)}{8 + 2m}}n^{\frac{m - 20}{32 + 8m}} \wedge c_4.
+	C_0\biggl(\frac{\log n}{n}\biggr)^{\frac{1}{m}} \leq r \leq c_0 \wedge M^{\frac{(m - 8)}{8 + 2m}} n^{\frac{m - 20}{32 + 8m}}.
 	\end{equation*}
@@
 	\label{thm:laplacian_smoothing_testing_manifold}
-	Let $f_0 \in H^1(\Xset,M)$ for $M \leq M_{\max}(m)$ and $m < 4$.
-Suppose that the neighborhood graph $G_{n,r}$ is computed with radius $r$ which satisfies~\ref{asmp:ls_kernel_radius_testing_manifold}, and the Laplacian smoothing test $\wh{\varphi}$ with $\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$, and $b \geq 1$.
-If
+	As in Theorem \ref{thm:laplacian_smoothing_testing}, but where $\Xset \subseteq \Rd$ is a manifold with intrinsic dimension $m < 4$, $M \leq M_{\max}(m)$, and the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}.
+	There are constants $N,C,c>0$ such that for any $n \geq N$, and any $r$ as in \ref{asmp:ls_kernel_radius_testing_manifold}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$}, satisfies the following: for all $b \geq 1$, and $f_0$ such that
 	\begin{equation}
@@
 	\end{equation} 
-	the Type II error is upper bounded:
-	\begin{align*}
-	& \Ebb_{f_0}\Bigl[1 - \wh{\varphi} \Bigr] \leq \\
-	& C\biggl(\frac{1}{b}\Bigl[1 + n^{-m/(4 + m)}M^{-2m/(4 + m)}\Bigr] + n\exp\bigl(-cnr^m\bigr)\biggr).
-	\end{align*}
+	the Type II error is upper bounded by
+	\begin{equation*}
+	\Ebb_{f_0}\bigl[1 - \wh{\varphi} \bigr] \leq C\biggl(\frac{1}{b}\Bigl[1 + n^{-m/(4 + m)}M^{-2m/(4 + m)}\Bigr] + n\exp\bigl(-cnr^m\bigr)\biggr).
+	\end{equation*}
+	% RJT: Alden, please verify that what I wrote is correct.
+	% AJG: Verified.
 \end{theorem}
-The proof of Theorems~\ref{thm:laplacian_smoothing_estimation_manifold} and~\ref{thm:laplacian_smoothing_testing_manifold} proceeds in a similar manner to that of Theorems~\ref{thm:laplacian_smoothing_estimation1} and~\ref{thm:laplacian_smoothing_testing}.
-The key difference is that in the manifold setting, the equations~\eqref{eqn:graph_sobolev_seminorm} and~\eqref{eqn:neighborhood_eigenvalue} that we use to upper bound bias and variance will hold with $d$ replaced by $m$.
 
-We emphasize that very little about $\Xset$ need be known for Theorems~\ref{thm:laplacian_smoothing_estimation_manifold} and~\ref{thm:laplacian_smoothing_testing_manifold} to hold.
-Indeed, all that is needed is the intrinsic dimension $m$--- in order to properly tune $r$ and $\rho$--- and otherwise $\wh{f}$ and $\wh{\varphi}$ are computed without regard to $\Xset$.
-In contrast, the penalty in~\eqref{eqn:thin_plate_spline} would have to be specially tailored to work in this setting, revealing another advantage of Laplacian smoothing compared to thin-plate splines.
+The proof of Theorems \ref{thm:laplacian_smoothing_estimation_manifold} and \ref{thm:laplacian_smoothing_testing_manifold} proceeds in a similar manner to that of Theorems \ref{thm:laplacian_smoothing_estimation1} and \ref{thm:laplacian_smoothing_testing}.
+The key difference is that in the manifold setting, the equations \eqref{eqn:graph_sobolev_seminorm} and \eqref{eqn:neighborhood_eigenvalue} used to upper bound bias and variance will hold with $d$ replaced by $m$.
+
+We emphasize that little about $\Xset$ needs be known for Theorems~\ref{thm:laplacian_smoothing_estimation_manifold} and \ref{thm:laplacian_smoothing_testing_manifold} to hold.
+Indeed, all that is needed is the intrinsic dimension $m$, to properly tune $r$ and $\rho$ (from a theoretical point of view), and otherwise \smash{$\wh{f}$} and \smash{$\wh{\varphi}$} are computed without regard to $\Xset$.
+In contrast, the penalty in \eqref{eqn:thin_plate_spline} would have to be specially tailored to work in this setting, revealing another advantage of the discrete approach to the variational one.
 
@@
 \label{sec:discussion}
-In this work, we have shown that Laplacian smoothing over a neighborhood graph is optimal for both estimation and goodness-of-fit testing problems.
 
-There are still several extensions we feel are worth pursuing.
-In practice, it is more common to use $k$NN graphs than neighborhood graphs, due to the guaranteed connectivity and sparsity of the former; we suspect that by building on the work of~\citet{calder2019}, one can show that our main results all hold when the $k$NN graph is used.
-In another direction, one can also generalize Laplacian smoothing by replacing the penalty $f^T \Lap_{n,r} f$ with $f^T \Lap_{n,r}^k f$ for some $k > 1$.
-The hope is that this adapted procedure would be minimax optimal over the previously mentioned higher-order Sobolev classes $H^k(\Xset)$.
-In the very special case of $k = 2$ and $d \leq 2$, we can show that this is indeed the case, but the general story--for all combinations of $k$ and $d$---remains beyond our reach.
+We have shown that Laplacian smoothing, computed over a neighborhood graph, can be optimal for both estimation and goodness-of-fit testing over Sobolev spaces.
+There are many extensions worth pursuing, and several have already been mentioned.
+We conclude by mentioning a couple more.
+In practice, it is more common to use a $k$-nearest-neighbor (kNN) graph than a neighborhood graph, due to the guaranteed connectivity and sparsity of the former; we suspect that by building on the work of \citet{calder2019}, one can show that our main results all hold under the kNN graph as well.
+In another direction, one can also generalize Laplacian smoothing by replacing the penalty \smash{$f^\top \Lap_{n,r} f$} with \smash{$f^\top \Lap_{n,r}^k f$}, for an integer $k > 1$.
+The hope is that this would then achieve minimax optimal rates over the higher-order Sobolev class $H^k(\Xset)$.
+% In the very special case of $k = 2$ and $d \leq 2$, we can show that this is indeed the case, but the general story--for all combinations of $k$ and $d$---remains beyond our reach.
+
+\subsection*{Acknowledgements}
+
+AG and RJT were supported by ONR grant N00014-20-1-2787.
 

### 2021-03-04 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
pg 1-6

@@
 
+\newcommand{\red}[1]{\textcolor{red}{#1}}
+\newcommand{\blue}[1]{\textcolor{blue}{#1}}
+\newcommand{\green}[1]{\textcolor{green}{#1}}
+\newcommand{\sbcomment}[1]{{\bf{{\red{{SB --- #1}}}}}}
+
 \usepackage{natbib}
-\renewcommand{\bibname}{REFERENCES}
+\renewcommand{\bibname}{References}
 \renewcommand{\bibsection}{\subsubsection*{\bibname}}
@@
 
-\begin{center} {\Large{\bf{Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs}}}
+\begin{center} {\Large{\bf{Minimax Optimal Regression over Sobolev Spaces \\
+\vspace{.2cm}
+ via Laplacian Regularization on Neighborhood Graphs}}}
 
@@
 
-\section{INTRODUCTION}
+\section{Introduction}
 
-We adopt the standard setup in nonparametric regression, where we observe samples $(X_1,Y_1),\ldots,(X_n,Y_n)$ that are i.i.d.\ draws from the model 
+We adopt the standard nonparametric regression setup, where we observe samples $(X_1,Y_1),\ldots,(X_n,Y_n)$ that are i.i.d.\ draws from the model 
 \begin{equation}
@@
 where $\varepsilon_i$ is independent of $X_i$.
-Our goal is to perform statistical inference on the unknown regression function $f_0$, by which we mean either \emph{estimating} $f_0$ or more simply \emph{testing} whether $f_0 = 0$, i.e., whether there is any signal present.
+Our goal is to perform statistical inference on the unknown regression function $f_0$, by which we mean either \emph{estimating} $f_0$ or \emph{testing} whether $f_0 = 0$, i.e., whether there is any signal present.
 
 Laplacian smoothing \citep{smola2003} is a penalized least squares estimator, defined over a graph.
-Letting $G = (V,W)$ be a weighted undirected graph with vertices $V=\{1,\ldots,n\}$, associated with $\{X_1,\ldots,X_n\}$, the Laplacian smoothing estimator \smash{$\wh{f}$} is given by
+Letting $G = (V,W)$ be a weighted undirected graph with vertices $V=\{1,\ldots,n\}$, associated with $\{X_1,\ldots,X_n\}$, and $W \in \mathbb{R}^{n \times n}$ is the (weighted) adjacency matrix of the graph.
+the Laplacian smoothing estimator \smash{$\wh{f}$} is given by
 \begin{equation}
@@
 \end{equation}
-Here $\Lap$ is the graph Laplacian matrix (to be defined in Section~\ref{sec:problem_setup_and_background}), $G$ is typically a geometric graph (such as a $k$-nearest-neighbor or neighborhood graph), $\rho \geq 0$ is a tuning parameter, and the penalty
+Here $\Lap$ is the graph Laplacian matrix (defined formally in Section~\ref{sec:problem_setup_and_background}), $G$ is typically a geometric graph (such as a $k$-nearest-neighbor or neighborhood graph), $\rho \geq 0$ is a tuning parameter, and the penalty
 \begin{equation*}
@@
 \end{equation}
-would in turn be a reasonable way to test $f_0 = 0$.
+is in turn a natural test statistic to test if $f_0 = 0$.
 
@@
 The upside of this work is its generality: in this setting $G$ need not be a geometric graph, but in principle it could be any graph over $V=\{1,\ldots,n\}$.
-The downside is that, in the context of nonparametric regression, it is not as natural to think of the evaluations of $f_0$ as exhibiting smoothness over some fixed pre-defined graph $G$, and more natural to speak of the smoothness of the function $f_0$ itself.
+The downside is that, in the context of nonparametric regression, it is arguably not as natural to think of the evaluations of $f_0$ as exhibiting smoothness over some fixed pre-defined graph $G$, and more natural to speak of the smoothness of the function $f_0$ itself.
 	\item \emph{Random design perspective.} Here one treats the design points $X_1,\ldots,X_n$ as independent samples from some distribution $P$ supported on a domain $\Xset \subseteq \Rd$.
@@
 In this paper, we will adopt the random design perspective, and seek to answer the following question:
-\begin{quote}{When we assume the regression function $f_0$ to be smooth in a continuum sense, does Laplacian smoothing achieve optimal performance for estimation and goodness-of-fit testing?} 
+\begin{quote}{When we assume the regression function $f_0$ is smooth in a continuum sense, does Laplacian smoothing achieve optimal performance for estimation and goodness-of-fit testing?} 
 \end{quote}
@@
 
-\section{SUMMARY OF RESULTS}
+\section{Summary of Results}
 
@@
 
-In light of this last point, it is worth asking whether there is any \emph{statistical} advantage to solving a discrete problem such as \eqref{eqn:laplacian_smoothing} (setting aside computational considerations for the moment).
+In light of the potential difficulty in bridging the gap between discrete and continuum notions of smoothness, it is worth asking whether there is any \emph{statistical} advantage to solving a discrete problem such as \eqref{eqn:laplacian_smoothing} (setting aside computational considerations for the moment).
 After all, we could have instead solved the following variational problem:
@@
 \label{eqn:thin_plate_spline_test}
-\wt{T} = \| \wt{f} \|_n^2 = \frac{1}{n} \sum_{i=1}^n \wt{f}(X_i)^2.
+\wt{T} = \| \wt{f} \|_n^2 := \frac{1}{n} \sum_{i=1}^n \wt{f}(X_i)^2.
 \end{equation}
 The penalty term in \eqref{eqn:thin_plate_spline} leverages the assumption that $f_0$ has a smooth derivative in a seemingly natural way.
-Indeed, the estimator \smash{$\wt{f}$} and statistic \smash{$\wt{T}$} are well-known: for $d = 1$, \smash{$\wt{f}$} is the familiar \emph{smoothing spline}, and for $d>1$, it is type of a \emph{thin-plate spline}.
+Indeed, the estimator \smash{$\wt{f}$} and statistic \smash{$\wt{T}$} are well-known: for $d = 1$, \smash{$\wt{f}$} is the familiar \emph{smoothing spline}, and for $d>1$, it is a type of \emph{thin-plate spline}.
 The statistical properties of smoothing and thin-plate splines are well-understood \citep{vandergeer2000,liu2019}.
@@
 Lastly, we use $a \wedge b = \max\{a,b\}$.
+\sbcomment{This is wrong, should check where it appears.}
 
-\section{BACKGROUND}
+\section{Background}
 \label{sec:problem_setup_and_background}
@@
 %we will make reference to these objects later on.
+\sbcomment{Don't understand why you're switching between $L$ and $L_{n,r}$?}
 
@@
 It's fine, I wouldn't change anything because it would involve explaining a lot more, I'm just pointing this out.
-(Throughout we assume $M \geq n^{-1/2}$, as otherwise the minimax rate is just the parametric rate $n^{-1}$.)
+(Throughout we assume $M \geq n^{-1/2}$, as otherwise the minimax rate is just the parametric rate $n^{-1}$.) \sbcomment{I don't think the second part of this comment is correct the way you've defined the Sobolev class.}
 
-As nonparametric hypothesis testing is (comparatively) less familiar than nonparametric estimation, we briefly summarize the main idea before stating the optimal error rate.
+As minimax rates in nonparametric hypothesis testing are (comparatively) less familiar than those in nonparametric estimation, we briefly summarize the main idea before stating the optimal error rate.
 In the goodness-of-fit testing problem, we ask for a test function---formally, a Borel measurable function $\phi$ taking values in $\{0,1\}$---which can distinguish between the hypotheses
@@
 In this paper, as in \citet{ingster2009}, we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\star} = 0$ and $\mc{F} = H^1(\Xset,M)$ is a first-order Sobolev ball.
+\sbcomment{Why isn't this without loss of generality?}
 
-The minimax critical radius is the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{F}_{\epsilon} := \mc{F} \cap \{f: \|f\|_{\Leb^2(\Xset)} \geq \epsilon\}$:
+\sbcomment{Define a level-$\alpha$ test.} The minimax critical radius is the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{F}_{\epsilon} := \mc{F} \cap \{f: \|f\|_{\Leb^2(\Xset)} \geq \epsilon\}$:
 \begin{equation*}
@@
 
-\section{MINIMAX OPTIMALITY OF LAPLACIAN SMOOTHING}
+\section{Minimax Optimality of Laplacian Smoothing}
 \label{sec:minimax_optimal_laplacian_smoothing}
@@
 
-\section{MANIFOLD ADAPTIVITY}
+\section{Manifold Adaptivity}
 \label{sec:manifold_adaptivity}
@@
 
-\section{DISCUSSION}
+\section{Discussion}
 \label{sec:discussion}
@@
 \bibliography{../../../graph_regression_bibliography}
-\clearpage
 

### 2021-03-04 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
pg 1-8

@@
 
-\sbcomment{Define a level-$\alpha$ test.} The minimax critical radius is the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{F}_{\epsilon} := \mc{F} \cap \{f: \|f\|_{\Leb^2(\Xset)} \geq \epsilon\}$:
+\sbcomment{Define a level-$\alpha$ test.
+You use it in~\eqref{eqn:type_I_error} for example.
+I also wouldn't re-use $\alpha$ for both Type-I and Type-II error.
+Later on you seem to use $b$ for Type-I error.
+Maybe use $b$ for Type-II and $\alpha$ for Type-I.} The minimax critical radius is the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{F}_{\epsilon} := \mc{F} \cap \{f: \|f\|_{\Leb^2(\Xset)} \geq \epsilon\}$:
 \begin{equation*}
@@
 	Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ has dimension $d < 4$ and $M \leq n^{1/d}$.
-Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume the neighborhood graph \smash{$G_{n,r}$} is computed with a kernel $K$ satifsying \ref{asmp:kernel}.
-There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_estimation}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$}, satisfies
+Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume the neighborhood graph \smash{$G_{n,r}$} is computed with a kernel $K$ satisfying \ref{asmp:kernel}.
+There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_estimation}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$},
+	with probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c (M^2n)^{d/(2 + d)})$, satisfies
 	\begin{equation*}
-	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(2 + d)} n^{-2/(2 + d)},
+	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(2 + d)} n^{-2/(2 + d)}.
 	\end{equation*}
-	with probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c (M^2n)^{d/(2 + d)})$.
 	% RJT: Alden, please verify that what I wrote is correct (I reorganized the way we explained this, added for $n \geq N$ ...).
@@
 For this reason, one really cannot speak of pointwise evaluation of a Sobolev function $f_0 \in H^1(\Xset)$ when $d>1$ (as we do in Theorem~\ref{thm:laplacian_smoothing_estimation1} by defining our target of estimation to be $f_0(X_i)$, $i=1,\ldots,n$).
-We can resolve this by appealing to what are known as \emph{Lebesgue points}, as explained in the appendix.
+We can resolve this by appealing to what are known as \emph{Lebesgue points}, as explained in Appendix~\ref{app:preliminaries}.
 	% RJT: Alden, please add a section to the supplement to explain this part.
@@
 	\item The lower bound on $r$ imposed in condition \ref{asmp:ls_kernel_radius_estimation} is compatible with practice, where by far the most common choice of radius is the connectivity threshold \smash{$r \asymp (\log(n)/n)^{1/d}$}, which makes \smash{$G_{n,r}$} as sparse as possible while still being connected, for maximum computational efficiency.
-The upper bound may seem a bit more mysterious---we need it for technical reasons to ensure that \smash{$\wh{f}$} does not overfit, but note that as a practical matter one rarely chooses $r$ to be so large anyway.
+The upper bound may seem a bit more mysterious---we need it for technical reasons to ensure that \smash{$\wh{f}$} does not overfit, but we note that as a practical matter one rarely chooses $r$ to be so large anyway.
 	\item It is possible to extend \smash{$\wh{f}$} to be defined on all of $\Xset$ and then evaluate the error of such an extension (as measured against $f_0$) in $\Leb^2(\Xset)$ norm.
@@
 We leave analysis of the Sobolev case to future work.
-	\item When $f_0$ is Lipschitz smooth, we can also replace the factor of $\delta$ in the high probability bound by the sharper $\delta^2/n$.
+	\item When $f_0$ is Lipschitz smooth, we can also replace the factor of $\delta$ in the high probability bound by the sharper $\delta^2/n$. \sbcomment{Too mysterious a comment. Why is this sharper?}
 \end{itemize}
@@
 When $d = 4$, our analysis results in an upper bound for the error of Laplacian smoothing that is within a \smash{$(\log n)^{1/3}$} factor of the minimax error rate.
-But when $d \geq 5$, our upper bounds are much worse.
+But when $d \geq 5$, our upper bounds do not match the minimax rates.
 \begin{theorem}
@@
 	\begin{equation*}
-	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{4/3} \biggl(\frac{\log n}{n}\biggr)^{-4/(3d)},
+	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{4/3} \biggl(\frac{\log n}{n}\biggr)^{4/(3d)},
 	\end{equation*}
@@
 \end{theorem}
+\sbcomment{I removed some negative sign above. Pls check.}
 This mirrors the conclusions of \citet{sadhanala16} who investigate estimation rates of Laplacian smoothing over the $d$-dimensional grid graph.

### 2021-03-04 manuscripts/laplacian_eigenmaps/appendix.tex
Completed section on manifold adaptivity of Laplacian Eigenmaps.

\noindent

\section{Proofs in Manifold Setting}
\label{sec:manifold_proofs}

In more detail, recall from our discuss.on in Section~\ref{subsec:analysis} that an essential step in our proofs is relating the iterated non-local operator $L_{P,\varepsilon}^j$ to the weighted Laplace-Beltrami operator $\Delta_P^j$, for $j = \floor{s/2}$. Suppose for the moment $s = 4$. Then we relate $L_{P,\varepsilon}$ to $\Delta_P$ by means of an intermediary nonlocal operator, as described by the following chain of estimates:
	\begin{align*}
	L_{P,\varepsilon} f(x) & = \int_{\mc{X}} \bigl(f(z) - f(x)\bigr)\eta\biggl(\frac{|z - x|}{\varepsilon}\biggr) \,dP(x) \\
	& = \bigl(1 + O(\varepsilon^2)\bigr) \int_{\mc{X}} \bigl(f(z) - f(z)\bigr)\eta\biggl(\frac{d_{\mc{X}}(z,x)}{\varepsilon}\biggr) \,dP(x) \\
	& = \Delta_Pf(x) + O\bigl(\varepsilon^2M\bigr) + O(\varepsilon^3).
	\end{align*}
	Here, $d_{\mc{X}}(\cdot,\cdot)$ is the geodesic distance on $\mc{X}$. For simplicity we have assumed $f \in C^3(\mc{X})$ and $p \in C^2(\mc{X})$, but the main idea will not change if we assume Sobolev smoothness instead. From this, we may deduce that
	\begin{align*}
	L_{P,\varepsilon}^2 f(x) = 
	\end{align*}
	When $s = 1$, this suffices.

### 2021-03-04 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Completed section on manifold adaptivity of Laplacian Eigenmaps.

@@
 	\item Regarding the restrictions on hyperparameters: the lower bound on $\varepsilon$ is the connectivity threshold, the smallest length scale at which the resulting graph will still, with high probability, be connected.
-On the other hand, as we will see in Section~\ref{sec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K \asymp \lambda_K(\Delta_P) \asymp K^{2/d}$.
+On the other hand, as we will see in Section~\ref{subsec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K \asymp \lambda_K(\Delta_P) \asymp K^{2/d}$.
 Finally, the restriction $K \asymp (M^2n)^{2d/(2 + d)}$ is chosen  to optimally trade-off bias and variance.
@@
 \subsection{Analysis}
-\label{sec:analysis}
+\label{subsec:analysis}
 
@@
 \begin{proposition}
+	\label{prop:graph_eigenvalue}
 	\textcolor{red}{(TODO)}
@@
 	\label{thm:laplacian_eigenmaps_estimation_manifold}
-	Suppose $f_0 \in H^s(\mc{X},M)$, for $s \leq 2$. Under appropriate conditions, with probability at least $1 - \delta - \textcolor{red}{(?)}$,
+	Suppose $f_0 \in H^s(\mc{X},M)$, for $s \leq 4$. Under appropriate conditions, with probability at least $1 - \delta - \textcolor{red}{(?)}$,
 	\begin{equation}
@@
 \begin{itemize}
-	\item \textbf{Proof strategy.} The proofs of Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold} proceed in a similar manner to the full-dimensional setting.
-	\item \textbf{Restriction on number of derivatives.} Unlike in the full-dimensional case, Theorem~\ref{thm:laplacian_eigenmaps_estimation_manifold} holds only when $s \leq 2$.
-Roughly speaking, we need this restriction in order to show that $L_{P,\varepsilon}^sf(x)$ is a good approximation to $\Delta_P^sf(x)$.
-Otherwise, the error incurred by using Euclidean distance rather than geodesic distance in~\eqref{eqn:nonlocal_laplacian} will dominate.
+	\item \textbf{Proof strategy.} The proofs of Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold} follow very similarly to the full-dimensional setting.
+The difference is that when $\mc{X}$ is a manifold with intrinsic dimension $m$, we can prove analogous results to Propositions~\ref{prop:graph_seminorm} and~\ref{prop:graph_eigenvalue}, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
+	\item \textbf{Restriction on number of derivatives.} Unlike in the full-dimensional case, our upper bound on the estimation error of $\wh{f}$ matches the minimax rate $n^{-2s/(2s + m)}$ only when $s \leq 4$.
+At a high level, thinking of the graph $G$ as an estimate of the manifold $\mc{X}$, we incur some error by using Euclidean distance rather than geodesic distance to form the edges of $G$.
+By contrast, in the full-dimensional setting the Euclidean metric exactly coincides with the geodesic distance for all points $x,z \in \mc{X}$ that are sufficiently close to each other and far from the boundary of $\mc{X}$).
+This extra error incurred in the manifold setting dominates when $s > 4$.
+We give a more detailed explanation of this phenomenon in Appendix~\ref{sec:manifold_proofs}.
+	\item \textbf{Comparison with traditional spectral series estimators.} Traditional spectral series estimators achieve the minimax rate for all values of $s$ and $m$.
+On the other hand, as we have already discussed, their default implementation assumes a priori knowledge of the domain $\mc{X}$, which is arguably a particularly unrealistic assumption when $\mc{X}$ is assumed to be a manifold.
+It is not clear whether this gap between spectral series and Laplacian eigenmaps estimators---or more generally, between estimators which assume strong knowledge of the manifold, and those which do not---is real, or a product of loose upper bounds.
 \end{itemize}
@@
 
+\appendix
+
+\input{appendix.tex}
+
 \end{document}

### 2021-03-04 notes/weekly_notes/12.24.20.NotesForWeek.tex
Completed section on manifold adaptivity of Laplacian Eigenmaps.

@@
 \section{Miscellaneous}
-I put things here which did not belong in any other section.
+I put things here which do not belong in any other section.
 
@@
 \end{equation*}
-We begin with a pointwise analysis of $L_{p,\varepsilon}u(x)$ at a given $x \in \Omega$.
+Let $\Omega_{\varepsilon} := \{x: B(x,\varepsilon) \in \Omega\}$ and let $\partial_{\varepsilon}\Omega := \Omega \setminus \Omega_{\varepsilon}$. Clearly
+\begin{equation*}
+\|L_{P,\varepsilon}u\|_{P}^2 = \underbrace{\int_{\Omega_{\varepsilon}} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_1} + \underbrace{\int_{\partial_{\varepsilon}\Omega} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_2}
+\end{equation*}
+and we now separately upper bound $I_1$ and $I_2$.
+
+\paragraph{Upper bound on $I_1$.}
+We begin with a pointwise analysis of $L_{p,\varepsilon}u(x)$ at a given $x \in \Omega_{\varepsilon}$.
 We may assume without loss of generality that $u \in C_0^{\infty}(\Omega)$, since $C_0^{\infty}(\Omega)$ is dense in $H_0^k(\Omega)$ and $\|T_{P,\varepsilon}u - u\|_P^2$ is continuous with respect to $\|\cdot\|_{H_0^k}$.
@@
 \end{equation}
-The way in which we use~\eqref{pf:kernel_smoothing_bias_1d_1} will depend on whether $x \in \Omega$ satisfies $B(x,\varepsilon) \subset \Omega$ or not.
-Let $\Omega_{\varepsilon} := \{x: B(x,\varepsilon) \in \Omega\}$ and let $\partial_{\varepsilon}\Omega := \Omega \setminus \Omega_{\varepsilon}$.
-Clearly
-\begin{equation*}
-\|L_{P,\varepsilon}u\|_{P}^2 = \underbrace{\int_{\Omega_{\varepsilon}} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_1} + \underbrace{\int_{\partial_{\varepsilon}\Omega} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_2}
-\end{equation*}
-and we now separately upper bound $I_1$ and $I_2$.
 
-\paragraph{Upper bound on $I_1$.}
-If $x \in \Omega_{\varepsilon}$ then all $z \in B(x,\varepsilon)$ also belong to $\Omega$. For one thing, this implies
+Now, since $x \in \Omega_{\varepsilon}$ then all $z \in B(x,\varepsilon)$ also belong to $\Omega$. For one thing, this implies
 \begin{equation*}
@@
 \end{equation*}
-Addtionally, we have that all $j = 1,\ldots,k - 1$, 
+Additionally, we have that all $j = 1,\ldots,k - 1$, 
 \begin{equation*}
@@
 \paragraph{Upper bound on $I_2$.}
+We will prove the upper bound
+\begin{equation*}
+\int_{0}^{\varepsilon} \bigl|L_{P,\varepsilon}u(x)\bigr|^2 \,dx \leq C \varepsilon^{2k},
+\end{equation*}
+with the bound on $\int_{1 - \varepsilon}^{1} \bigl|L_{P,\varepsilon}u\bigr|^2$ following in a completely analogous fashion.
+
+Taking a Taylor expansion of $u$ around $0$ gives
+\begin{align*}
+u(x) - u(0) & = \sum_{j = 1}^{k - 1} x^j u^{(j)}(0) + \frac{1}{(k - 1)!} \int_{0}^{x}u^{(k)}(s)(x - s)^{k - 1}\,ds \\
+& = \frac{1}{(k - 1)!} \int_{0}^{x}u^{(k)}(s)(x - s)^{k - 1}\,ds,
+\end{align*}
+with the second inequality following since $u \in C_c^{\infty}(\Omega)$ implies $u^{(j)}(0) = 0$ for $j = 1,\ldots,k - 1$. From this identity along with Jensen's inequality, we have
+\begin{align*}
+\int_{0}^{\varepsilon} \int_{\Omega} \bigl(u(x)\bigr)^2 \eta\biggl(\frac{z - x}{\varepsilon}\biggr) \,dz \,dx & = \int_{0}^{\varepsilon} d_{\nu}(x) \bigl(u(x)\bigr)^2 \,dx \\
+& \leq \varepsilon \int_{0}^{\varepsilon} \bigl(u(x)\bigr)^2 \,dx \\
+& \leq \varepsilon^{2k} \int_{0}^{\varepsilon} \int_{0}^{x} \bigl(u^{(k)}(s)\bigr)^2 \,ds \,dx. 
+\end{align*}
+From here, switching the order of integration and substituting $v = s/\varepsilon$ gives the desired result:
+\begin{align*}
+I_1 & \leq \frac{\varepsilon^{2k - 1}}{\alpha} \int_{0}^{\varepsilon} \int_{s}^{\varepsilon} \bigl(u^{(k)}(s)\bigr)^2 \,dx \,ds \\
+& \leq \frac{\varepsilon^{2k + 1}}{\alpha} \int_{0}^{1} \bigl(u^{(k)}(v)\bigr)^2 \,dv \\
+& = \frac{\varepsilon^{2k + 1}}{\alpha} |u|_{H^k(\Omega)}^2
+\end{align*}
 

### 2021-03-07 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
till just before B.2

@@
 \end{enumerate}
-We will also require that the radius of the graph Sobolev class not be too large. Precisely, we will require $M \leq M_{\max}(d)$, where we define
+We will also require that the radius of the graph Sobolev class \sbcomment{What does the class have to do with graphs?} not be too large.
+Precisely, we will require $M \leq M_{\max}(d)$, where we define
 \begin{equation*}
@@
 \end{theorem}
+\sbcomment{Personally, I wouldn't emphasize the Type II error so much in the Theorem (it is not tight and it's not an interesting quantity anyway), and write \eqref{eqn:laplacian_smoothing_testing} more directly as an upper bound on the critical radius (which is the interesting quantity).
+Otherwise you've introduced that terminology for no reason.}
 
@@
 On the other hand, if we explicitly assume that $f_0 \in \Leb^4(\Xset,M)$, then \citet{guerre02} show that the testing problem is characterized by a dimension-free lower bound $\epsilon^{2}(\Leb^4(\Xset,M)) \gtrsim n^{-1/2}$.
-Moreover, by training \smash{$\wh{f}$} to the interpolation limit, setting $\rho = 0$, the subsequent test \smash{$\wh{\varphi}$} will achieve (up to constants) this lower bound.
+Moreover, by estimating \smash{$\wh{f}$}, setting $\rho = 0$, the subsequent test \smash{$\wh{\varphi}$} will achieve (up to constants) this lower bound.
 That is, for any $f_0 \in \Leb^4(\Xset,M)$ such that \smash{$\|f_0\|_{\Leb^2(\Xset)}^2 \geq C b^2n^{-1/2}$}, we have both
@@
 \end{equation*}
-and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies e.g., Dirichlet boundary conditions.} To be clear, this argument does not formally imply that $\wh{f}$ and $\wt{f}$ are close (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of continuous functions with weak derivatives in $\Leb^2(\Xset)$).
+and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies e.g., Dirichlet boundary conditions.} To be clear, this argument does not formally imply that $\wh{f}$ and $\wt{f}$ are close \sbcomment{Should remind people what these are, or refer to eqns where they are defined.} (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of continuous functions with weak derivatives in $\Leb^2(\Xset)$).
 But it does seem to suggest that the two estimators should behave somewhat similarly.
@@
 While these solutions do improve the statistical properties of \smash{$\wt{f}$} for $d > 1$ (see e.g., \citet{birge1993,birge1998,vandergeer2000}), Laplacian smoothing is generally speaking much simpler and more computationally friendly.
-In addition, the other approaches are specifically tailored to the domain $\Xset$, in stark contrast to \smash{$\wh{f}$}.
+In addition, the other approaches are usually specifically tailored to the domain $\Xset$, in stark contrast to~\smash{$\wh{f}$}.
 
@@
 The parameter $\rho$ is then tuned to minimize the sum of bias and variance, in the usual manner.
+\sbcomment{Should probably say something about the rest of the proof, i.e. is it straightforward to put these together to get the two Theorems?}
 
@@
 On the other hand, a theory has been developed \citep{belkin03,belkin05,niyogi2013} establishing that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
-We contribute to this line of work by showing that under the manifold hypothesis, Laplacian smoothing achieves the sharper minimax rates over $H^1(\Xset)$.
+We contribute to this line of work by showing that under the manifold hypothesis, Laplacian smoothing achieves the tighter minimax rates over $H^1(\Xset)$.
 
@@
 \end{theorem}
-
+In a similar vein, we obtain results for manifold adaptive testing under the following condition on the graph radius parameter:
 \begin{enumerate}[label=(R\arabic*)]
@@
 \end{theorem}
+\sbcomment{Same comment as before on what to emphasize.}
 
@@
 Indeed, all that is needed is the intrinsic dimension $m$, to properly tune $r$ and $\rho$ (from a theoretical point of view), and otherwise \smash{$\wh{f}$} and \smash{$\wh{\varphi}$} are computed without regard to $\Xset$.
-In contrast, the penalty in \eqref{eqn:thin_plate_spline} would have to be specially tailored to work in this setting, revealing another advantage of the discrete approach to the variational one.
+In contrast, the penalty in \eqref{eqn:thin_plate_spline} would have to be specially tailored to work in this setting, revealing another advantage of the discrete approach over the variational one.
 
@@
 % In the very special case of $k = 2$ and $d \leq 2$, we can show that this is indeed the case, but the general story--for all combinations of $k$ and $d$---remains beyond our reach.
+\sbcomment{Not a huge fan of using $k$ in adjacent lines to refer to completely different things.}
 
 \subsection*{Acknowledgements}
-
-AG and RJT were supported by ONR grant N00014-20-1-2787.
+AG and RJT were supported by ONR grant N00014-20-1-2787. AG and SB were supported by NSF grants DMS-1713003 and CCF-1763734.
 

### 2021-03-07 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Added first part of kernel smoothing bias.

@@
 \DeclareMathAccent{\wb}{0}{mathx}{"73}
+\DeclareMathAccent{\wc}{0}{mathx}{"71}
 
@@
 \textbf{(1) Geometric graphs.} In graph-based learning, one observes data $X_1,\ldots,X_n$ sampled independently from an unknown distribution $P$, and forms a geometric graph $G$---with edges $e_{ij}$ corresponding to proximity between samples $X_i$ and $X_j$---over the observed data.
-Geometric graphs encode information about $P$ in an extremely general manner, and can thus be leveraged to conduct many different fundamental statistical tasks.
+Geometric graphs encode information about $P$ in an extremely general manner, and can be leveraged to conduct many different fundamental statistical tasks.
 These include clustering, semi-supervised learning, classification, and regression---both estimation and goodness-of-fit testing.
@@
 
-\textbf{(2) What we study: Regression and Laplacian eigenmaps.} In this paper we focus on regression where in addition to the design points $X_1,\ldots,X_n$ one observes real-valued responses $Y_1,\ldots,Y_n$, and learns the regression function $f_0(x) := E[Y|X = x]$.
+\textbf{(2) What we study: Regression and Laplacian eigenmaps.} In this paper we focus on regression where in addition to the design points $X_1,\ldots,X_n$ one observes real-valued responses $Y_1,\ldots,Y_n$, and seeks to learn the unknown regression function $f_0(x) := E[Y|X = x]$.
 We consider both the estimation and goodness-of-fit testing problems.
 The methods we will study are based on \emph{Laplacian eigenmaps}, first introduced by \cite{belkin03a}, which projects the response vector $Y = (Y_1,\ldots,Y_n)$ onto the span of the leading eigenvectors of the graph Laplacian $L$.
-The Laplacian $L$ is a difference operator, acting on functions $f: {\bf X} \to \Reals$ as follows,
+The Laplacian $L$ is a difference operator, acting on functions $f: \{X_1,\ldots,X_n\} \to \Reals$ as follows,
 \begin{equation}
@@
 The eigenvectors of $L$ form an orthonormal basis of $L^2(X)$, and are approximations to the eigenvectors of $\Delta_P$.
-Their corresponding eigenvalues are estimates of~\textcolor{red}{(...)}, and provide a notion of smoothness for each eigenvector---the smaller the eigenvalue, the smoother the eigenvector.
+Their corresponding eigenvalues are estimates of~\textcolor{red}{(...)}, and provide a notion of smoothness for each eigenvector---the smaller the eigenvalue, the smoother the corresponding eigenvector.
 
@@
 We observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$, where the design points $X_1,\ldots,X_N$ are drawn i.i.d from a distribution $P$ supported on a compact set $\mc{X} \subseteq \Rd$, and responses
-\begin{equation*}
+\begin{equation}
+\label{eqn:model}
 Y_i = f_0(X_i) + \epsilon_i,
-\end{equation*}
+\end{equation}
 with regression function $f_0: \mc{X} \to \Reals$, and $\epsilon_i \sim N(0,1)$ independent Gaussian noise.
@@
 \end{equation*}
-The estimator $\wt{f}$ and a test based on $\wt{T}$ are minimax rate-optimal over $H_0^s(\mc{X})$, a fact intimately related to the series representation of periodic Sobolev spaces, e.g.~\eqref{eqn:periodic_sobolev_space}.
+The estimator $\wt{f}$ and a test based on $\wt{T}$ are minimax rate-optimal over $H_0^s(\mc{X})$.
+This fact is intimately related to the series representation of periodic Sobolev spaces, e.g.~\eqref{eqn:periodic_sobolev_space}, whereby the model~\eqref{eqn:model} can be formally reduced to the Gaussian sequence model (see e.g.
+\citep{tsybakov08,gine16}).
 
@@
 Typically one does not have direct knowledge of the design distribution $P$, and even if $P$ is known, the eigenfunctions of $\Delta_P$ are rarely easy to compute.
-As an alternative, suppose instead we had access only to eigenfunctions $f_1',f_2',\ldots$ of the unweighted Laplace-Beltrami operator $\Delta$: then, letting $F_K = \mathrm{span}\{f_1,\ldots,f_K\}$, the least-squares estimator and test statistic
+As an alternative, suppose instead we had access only to eigenfunctions $f_1',f_2',\ldots$ of the unweighted Laplace-Beltrami operator $\Delta_{\nu}$.
+Then letting $F_K = \mathrm{span}\{f_1,\ldots,f_K\}$, the least-squares estimator and test statistic
 \begin{equation*}
@@
 As this view suggests, Laplacian eigenmaps naturally incurs some additional error compared to traditional spectral series estimators---this is the price we pay for using an approximation of $\Delta_P$.
-Fortunately, despite this extra error Laplacian eigenmaps retains the favorable optimality properties of its classical spectral series counterparts while freeing us from needing unrealistic knowledge of $\Delta_P$, as now summarize.
+Fortunately, the general message of this paper is that despite this extra error, Laplacian eigenmaps retains many of the favorable optimality properties of its classical spectral series counterparts while freeing us from requiring unrealistic knowledge of $\Delta_P$.
 
@@
 	\item \textbf{Manifold adaptivity.}
-	If $\mc{X} \subset \Rd$ is a manifold of dimension $m < d$, and if $H^s(\mc{X})$ for $s \leq 2$, then each of the aforementioned rates hold with $d$ replaced by $m$.
+	If $\mc{X} \subset \Rd$ is a manifold of dimension $m < d$, and if $H^s(\mc{X})$ for $s \leq 4$, then each of the aforementioned rates hold with $d$ replaced by $m$.
 \end{itemize}
@@
 \subsection{Error rates under the manifold hypothesis}
-We now show that Laplacian eigenmaps achieves the faster minimax rates under the manifold hypothesis, when the regression function $f_0 \in H^1(P;M)$ or $f_0 \in H^2(P;M)$.
-This section will proceed in a similar fashion to~\ref{sec:higher_order_sobolev_classes}, except with the ambient dimension $d$ replaced by intrinsic dimension $m$.
+We now show that Laplacian eigenmaps achieves the faster minimax rates under the manifold hypothesis, when the regression function $f_0 \in H^s(P;M)$ and $s \leq 4$.
+This section will proceed in a similar fashion to Section~\ref{sec:higher_order_sobolev_classes}, except with the ambient dimension $d$ replaced by intrinsic dimension $m$.
 
@@
 	\item \textbf{Comparison with traditional spectral series estimators.} Traditional spectral series estimators achieve the minimax rate for all values of $s$ and $m$.
-On the other hand, as we have already discussed, their default implementation assumes a priori knowledge of the domain $\mc{X}$, which is arguably a particularly unrealistic assumption when $\mc{X}$ is assumed to be a manifold.
-It is not clear whether this gap between spectral series and Laplacian eigenmaps estimators---or more generally, between estimators which assume strong knowledge of the manifold, and those which do not---is real, or a product of loose upper bounds.
+On the other hand, as we have already discussed, these estimators assume a priori knowledge of the domain $\mc{X}$, which is arguably a particularly unrealistic assumption when $\mc{X}$ is assumed to be a manifold.
+It is not clear whether this gap between spectral series and Laplacian eigenmaps estimators---or more generally, between estimators which assume the manifold is known, and those which do not---is real, or a product of loose upper bounds.
+\end{itemize}
+
+\textcolor{red}{(TODO)}
+\begin{itemize}
+	\item \textcolor{red}{Give a reference regarding the optimality of spectral series estimators on known manifolds.}
 \end{itemize}
@@
 \label{sec:out_of_sample}
-Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity} show that Laplacian eigenmaps is an optimal method for nonparametric regression.
-As explained previously, we have measured estimation error in $\|\cdot\|_n$ norm---in words, the average squared loss incurred over the $n$ observed design points--- whereas error measured in $L^2(P)$ norm is the more typical metric in the random design setup.
+To summarize, Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity} show that $\wh{f}$ is a minimax optimal estimator over Sobolev spaces.
+However, as mentioned previously we have measured loss \emph{in-sample}---that is, measured in $\Leb^2(P_n)$ norm---whereas \emph{out-of-sample} error---error measured in $L^2(P)$ norm---is the more typical metric in the random design setup.
 
 Of course, the Laplacian eigenmaps estimator is only defined at the observed design points $X_1,\ldots,X_n$, and to measure its error in $L^2(P)$ norm we must first extend it to the rest of $\Xset$.
-We propose a simple method, essentially kernel smoothing, to do the job, which can be applied to any estimator defined at the design points, including Laplacian eigenmaps.
-We show that the smoothed version of $\wh{f}$ has optimal $L^2(P)$ error.
+We propose a simple method, kernel smoothing, to do the job.
+The method can applied to any estimator defined at the design points, including Laplacian eigenmaps, and we show that a smoothed version $\wt{f}$ of our original estimator $\wh{f}$ has optimal $L^2(P)$ error.
+For simplicity, in this section we stick to the full-dimensional setting, but everything can be adapted to the manifold setting in a straightforward way.
 
-\subsection{Extension using Kernel Smoothing}
+\paragraph{Extension by kernel smoothing.}
+We now formally define our approach to extending by kernel smoothing.
+For a kernel function $\psi(\cdot): [0,\infty) \to (-\infty,+\infty)$, bandwidth $h > 0$, and a distribution $Q$, the \emph{Nadaraya-Watson kernel smoother} $T_{\varepsilon,Q}$ is given by
+\begin{equation*}
+\bigl(T_{h,Q}f)(x) := \frac{1}{d_Q(x)} \int_{\Omega} f(z)\psi\biggl(\frac{\|z - x\|}{h}\biggr) \,dQ(z)
+\end{equation*}
+where $d_Q(x) := \int_{\Omega} \psi\bigl(\|z - x\|/\varepsilon\bigr) \,dQ(z)$. We denote $T_{\varepsilon,P_n}$ by $T_{\varepsilon,n}$, and $d_{P_n}$ by $d_n$.
 
-\subsection{Out-of-sample error of Kernel Smoothed Laplacian Eigenmaps}
+The estimated function $\wt{f}$ we will consider is a kernel smoother passed over $\wh{f}$, meaning
+\begin{equation}
+\label{eqn:kernel_smoother_laplacian_eigenmaps}
+\wt{f}(x) := \bigl(T_{h,n}\wh{f}\bigr)(x).
+\end{equation}
+It is useful to make a few comments about $\wt{f}$.
+First of all, while in principle~\eqref{eqn:kernel_smoother_laplacian_eigenmaps} makes sense for any $x \in \Rd$, we will always think of $\wt{f}$ as a function on $\mc{X}$.
+Additionally, note we may use the same kernel for smoothing as was used to construct the graph, i.e.
+take $\psi = \eta$.
+Likewise, we can set $h = \varepsilon$.
+However, we will see that for theoretical purposes these are not always the best choices.
 
-\begin{theorem}
-	\label{thm:kernel_smoothing}
-	\textcolor{red}{(TODO)}
-\end{theorem}
+\paragraph{Out-of-sample error of kernel smoothed Laplacian Eigenmaps.}
+
+In Lemma~\ref{lem:kernel_smoothing_laplacian_eigenmaps}, we consider an arbitrary estimator $\wc{f} \in L^2(P_n)$.
+We show that the out-of-sample error $\|T_{n,h}\wc{f} - f_0\|_P^2$ can be upper bounded by three terms--- (a constant times) the in-sample error of $\wc{f} - f_0$, and bias and variance terms associated with a noiseless version of kernel smoothing.
+We shall assume the following conditions on $\psi$.
+\begin{enumerate}[label=(K\arabic*)]
+	\setcounter{enumi}{1}
+	\item
+	\label{asmp:kernel}
+	The kernel function $\psi$ is supported on a subset of $[0,1]$, with $\psi(1/2) > 0$.
+Additionally, $\psi$ is Lipschitz continuous on $[0,1]$.
+Finally, for convenience, we will assume that $\psi$ is normalized such that
+	\begin{equation*}
+	\int_{-\infty}^{\infty} \psi(|s|) \,ds = 1.
+	\end{equation*}
+\end{enumerate}
+\begin{lemma}
+	\label{lem:kernel_smoothing_laplacian_eigenmaps}
+	Suppose $\wc{f} \in L^2(P_n)$, $f_0 \in H^1(\mc{X})$ and $p \in C^1(\mc{X})$. Under Assumptions~\ref{asmp:domain}, \ref{asmp:density}, and~\ref{asmp:kernel}, it holds that
+	\begin{equation}
+	\label{eqn:kernel_smoothing_laplacian_eigenmaps}
+	\|T_{n,h}\wc{f} - f_0\|_P^2 \leq C\biggl(\|\wc{f} - f_0\|_n^2 + \frac{1}{\delta} \cdot \frac{h^2}{nh^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{h,P}f_0 - f_0\|_P^2\biggr),
+	\end{equation}
+	with probability at least $1 - \delta - Ch^d\exp\{-Cnh^d\}$. 
+\end{lemma}
+Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data, by a factor of $h^2$.
+The bias term, on the other hand, is standard (at least when $f_0$ is assumed to be H\"{o}lder smooth), and can be tightly bounded using Taylor expansion when $\psi$ is an order-$s$ kernel.
+Recall that $\psi$ is an order-$s$ kernel if
+\begin{equation*}
+\int_{-\infty}^{\infty} \psi(|s|) = 1,~~\int_{-\infty}^{\infty} t^j \psi(|t|) \,dt = 0~~\textrm{for $j = 1,\ldots,s - 1$},~~\textrm{and}~~\int_{-\infty}^{\infty} t^s \psi(|t|) \,dt < \infty.
+\end{equation*}
+\begin{lemma}
+	\label{lem:kernel_smoothing_bias}
+	Suppose the domain $\mc{X}$, the distribution $P$, and the kernel $\psi$  satisfy~\ref{asmp:domain}, \ref{asmp:density}, and~\ref{asmp:kernel}, respectively. 
+	\begin{itemize}
+		\item If $f_0 \in H^1(P)$, it follows that
+		\begin{equation*}
+		\|T_{h,P}f_0 - f_0\|_P^2 \leq C h^{2} |f|_{H^1(P)}^2.
+		\end{equation*}
+		\item If $f_0 \in H_0^{s}(P)$, $p \in C^{r}(\mc{X})$ for $r = s - 1$, and $\psi$ is an order-$s$ kernel, it follows that
+		\begin{equation*}
+		\|T_{h,P}f_0 - f_0\|_P^2 \leq C h^{2s} |f|_{H^s(\mc{X})}^2.
+		\end{equation*}
+	\end{itemize}
+\end{lemma}
+Choosing $h \asymp n^{-1/(2r + d)}$ for $r = s - 1$ balances the kernel smoothing bias and variance terms in~\eqref{eqn:kernel_smoothing_laplacian_eigenmaps}.
+Then, applying the bounds in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} (when $s = 1$) or Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} (when $s > 1$), we conclude that $\wt{f}$ is an optimal estimator, error measured in $L^2(P)$ norm.
 
@@
 State explicitly that we choose kernel smoothing for its simplicity, ubiquity, and strong theoretical properties under our assumptions, but that considerations of extensions for specifically tailored to Laplacian based estimators could be of interest.
+	\item \textbf{Comparison with kernel smoothing.}
 \end{itemize}
@@
 \begin{itemize}
-	\item \textbf{()}
+	\item \textbf{Other properties of spectral series estimators.}
 \end{itemize}

### 2021-03-07 notes/weekly_notes/12.24.20.NotesForWeek.tex
Added first part of kernel smoothing bias.

--- 2021-03-03 notes/weekly_notes/12.24.20.NotesForWeek.tex
@@
 \end{equation*}
-where $d_Q(x) = \int_{\Omega} \eta\bigl(\|z - x\|/\varepsilon\bigr) \,dQ(z)$.
-We denote $T_{\varepsilon,P_n}$ by $T_{\varepsilon,n}$, and $d_{P_n}$ by $d_n$.
-For convenience, we will assume that $\eta$ is normalized, $\int_{0}^{\infty} \eta(s) \,ds = 1$, and we will denote $K(\cdot) = (\eta(\cdot))^2$.
+where $d_Q(x) = \int_{\Omega} \eta\bigl(\|z - x\|/\varepsilon\bigr) \,dQ(z)$. We denote $T_{\varepsilon,P_n}$ by $T_{\varepsilon,n}$, and $d_{P_n}$ by $d_n$.
 
@@
 	\label{asmp:kernel}
-	The kernel function $\eta$ is supported on a subset of $[0,1]$, with $\eta(1/2) > 0$. Additionally, $\eta$ is \textcolor{red}{Lipschitz} on its support.
+	The kernel function $\eta$ is compactly supported on a subset of $[0,1]$. Additionally, $\eta$ is Lipschitz on its support. Finally, for convenience we will assume that $\eta$ is normalized, 
+	\begin{equation*}
+	\int_{0}^{\infty} \eta(s) \,ds = 1.
+	\end{equation*}
 	\item 
@@
 Assumption~\ref{asmp:domain} is thus a condition on the boundary $\partial \Omega$, and is implied for all sufficiently small $\varepsilon$ if the boundary is Lipschitz.
+Denoting $K(\cdot) = (\eta(\cdot))^2$, we note that Assumption~\ref{asmp:kernel} implies that $\eta$ is square-integrable, $\int_{0}^{\infty} K(s) \,ds < \infty$.
 \begin{lemma}
@@
 \begin{proof}[Proof (when dimension $d = 1$).]
+	\textcolor{red}{(TODO)}
 \end{proof}
@@
 
-	\textit{Proof of \eqref{pf:kernel_smoothing_laplacian_eigenmaps_0}.} We will begin by assuming that $\eta(z) \geq 0$ for all $z$, and return to prove the claim in its full generality at the end.
+	\textit{Proof of \eqref{pf:kernel_smoothing_laplacian_eigenmaps_0}.} We will begin by assuming that $\eta$ is positive, $\eta(z) \geq 0$ for all $z$, and return to prove the claim in its full generality at the end.
 
-	The left hand side of~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is equal to $\|T_{\varepsilon,n}(\wh{\theta}_{1:n} - f^{\ast}_{1:n})\|_P$, and~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is implied by showing that the operator $T_{\varepsilon,n}: L^2(X) \to L^2(\Omega)$ is (up to a constant) a contraction, meaning that for any $u \in L^2(X)$,
+	The left hand side of~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is equal to $\|T_{\varepsilon,n}(\wh{\theta}_{1:n} - f^{\ast}_{1:n})\|_P$, and~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is implied by showing that the operator $T_{\varepsilon,n}: L^2(X) \to L^2(\Omega)$ is (up to a constant) a contraction.
+We will show that for any $u \in L^2(X)$,
 	\begin{equation}
@@
 	\begin{equation*}
-	\sup_{x \in \Omega}|d_n(x) - d_P(x)| = o_P(\varepsilon^d);
+	\sup_{x \in \Omega}|d_n(x) - d_P(x)| = o_P(\varepsilon^d).
 	\end{equation*}
-	moreover, it follows from Lipschitz continuity of $p$ that for any $x \in \Omega$,
+	Moreover, it follows from the Lipschitz continuity of $p$ that for any $x \in \Omega$,
 	\begin{equation*}
-	d_P(x) = \int_{\Omega} \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) p(x') \,dx' = p(x) d_{\nu}(x) + O(\varepsilon^{d + 1});
+	d_P(x) = \int_{\Omega} \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) p(x') \,dx' = p(x) d_{\nu}(x) + O(\varepsilon^{d + 1}),
 	\end{equation*}
@@
 \section{Miscellaneous}
-I put things here which did not belong in any other section.
+I put things here which do not belong in any other section.
 
 \subsection{Proof of Lemma~\ref{lem:kernel_smoothing_bias}, $d = 1$ and $p$ uniform}
+\label{subsec:kernel_smoothing_bias_pf_1d_uniform_p}
 Without loss of generality take the domain $\Omega = (0,1)$. Recall that we are interested in upper bounding
@@
 L_{P,\varepsilon}u(x) & = \frac{1}{\varepsilon} \cdot \frac{1}{(k - 1)!} \int_{0}^{1} \int_{\Omega} (1 - t)^k (z - x)^k u^{(k)}(x + t(z - x)) \eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dz \,dv \\
-& = \frac{\varepsilon^{k}}{(k - 1)!} \int_{0}^{1} \int_{B(0,1)} (1 - t)^k v^k u^{(k)}(x + t\varepsilon v) \eta(v) \,dv \,dt.
+& = \frac{\varepsilon^{k}}{(k - 1)!} \int_{0}^{1} \int_{B(0,1)} (1 - t)^k v^k u^{(k)}(x + t\varepsilon v) \eta(|v|) \,dv \,dt
 \end{align*} 
-This identity along with Jensen's inequality implies that
+and as a result
 \begin{equation*}
-\Bigl(L_{P,\varepsilon}u(x)\Bigr)^2 \leq \biggl(\frac{\varepsilon^{k}}{(k - 1)!}\biggr)^2 \cdot  \int_{0}^{1} \int_{B(0,1)} v^{2k} \bigl(u^{(k)}(x + t\varepsilon v)\bigr)^2 \eta(|v|) \,dv \,dt.
+\Bigl(L_{P,\varepsilon}u(x)\Bigr)^2 = \frac{\varepsilon^{2k}}{(k!)^2} {\sigma_\eta}^2 \bigl(u^{(k)}(x)\bigr)^2 + \frac{\varepsilon^{2k}}{((k - 1)!)^2}A_1(x)A_{2}(x),
 \end{equation*}
-Integrating over all $x \in \Omega_{\varepsilon}$ gives an upper bound on $I_1$. 
-From here, by changing the order of integration, substituting $\wt{x} = x + t\varepsilon v$, and noting that $\wt{x} \in \Omega$, we arrive at the desired bound:
+where $\sigma_\eta = \int v^k \eta(|v|) \,dv$ and 
 \begin{align*}
-I_1 & \leq \biggl(\frac{\varepsilon^{k}}{(k - 1)!}\biggr)^2 \int_{0}^{1} \int_{B(0,1)} \int_{\Omega_{\varepsilon}} v^{2k} \bigl(u^{(k)}(x + t\varepsilon v)\bigr)^2 \eta(|v|) \,dx \,dv \,dt \\
-& \leq \biggl(\frac{\varepsilon^{k}}{(k - 1)!}\biggr)^2 \int_{0}^{1} \int_{B(0,1)} \int_{\Omega} v^{2k} \bigl(u^{(k)}(\wt{x})\bigr)^2 \eta(|v|) \,d\wt{x} \,dv \,dt \\
-& = \biggl(\frac{\varepsilon^{k}}{(k - 1)!}\biggr)^2 |u|_{H^k(\Omega)}^2 \sigma_{\eta}.
+A_1(x) & := \int_{0}^{1} \int_{B(0,1)} (1 - t)^k v^k \bigl(u^{(k)}(x + t\varepsilon v) - u^{(k)}(x)\bigr) \eta(|v|) \,dv \,dt \\
+A_2(x) & := \int_{0}^{1} \int_{B(0,1)} (1 - t)^k v^k \bigl(u^{(k)}(x + t\varepsilon v) + u^{(k)}(x)\bigr) \eta(|v|) \,dv \,dt.
 \end{align*}
+Integrating over all $x \in \Omega_{\varepsilon}$, we have that
+\begin{align*}
+I_1 = \int_{\Omega_{\varepsilon}} \bigl(L_{P,\varepsilon}u(x)\bigr)^2 \,dx & = \frac{\varepsilon^{2k}}{(k!)^2} \sigma_\eta^2 \int_{\Omega_{\varepsilon}} \bigl(u^{(k)}(x)\bigr)^2 \,dx + \frac{\varepsilon^{2k}}{((k - 1)!)^2} \int_{\Omega_{\varepsilon}}A_1(x) A_2(x) \,dx \\
+& \leq \frac{\varepsilon^{2k}}{(k!)^2} \Bigl(\sigma_\eta^2 \|u^{(k)}\|_{P}^2 + k^2 \|A_1\|_P \|A_2\|_P\Bigr).
+\end{align*}
+It remains to show that $\|A_1\|_P\|A_2\|_P = o(1)$.
+
+\textcolor{red}{(TODO): Follow through on this calculation, referring to pages 192 - 195 of Tsybakov's book.}
 
 \paragraph{Upper bound on $I_2$.}
+We will prove the upper bound
+\begin{equation}
+\label{pf:kernel_smoothing_bias_1d_2}
+\int_{0}^{\varepsilon} \bigl|L_{P,\varepsilon}u(x)\bigr|^2 \,dx \leq C \varepsilon^{2k + 1}\|u^{(k)}\|_P^2,
+\end{equation}
+with the bound on $\int_{1 - \varepsilon}^{1} \bigl|L_{P,\varepsilon}u\bigr|^2$ following in a completely analogous fashion.
+
+Unlike the previous case, when $x$ is near the boundary of $\Omega$ we cannot use the higher-order property of $\eta$ to cancel the order-$1$ up through order-$(k - 1)$ derivatives of $u$.
+Instead, we will use the zero-trace property to show that $u$ itself must be quite small.
+Taking a Taylor expansion of $u$ around $0$, we have that for any $x \in \Omega$,
+\begin{align}
+u(x) & = u(0) + \sum_{j = 1}^{k - 1} x^j u^{(j)}(0) + \frac{1}{(k - 1)!} \int_{0}^{x}u^{(k)}(s)(x - s)^{k - 1}\,ds \nonumber \\
+& = \frac{1}{(k - 1)!} \int_{0}^{x}u^{(k)}(s)(x - s)^{k - 1}\,ds. \label{pf:kernel_smoothing_bias_1d_3}
+\end{align}
+with the second inequality following since $u \in C_c^{\infty}(\Omega)$ implies $u(0) = 0$ as well as $u^{(j)}(0) = 0$ for $j = 1,\ldots,k - 1$.
+To make use of this upper bound on $u(x)$, we upper bound $|L_{P,\varepsilon}u(x)|^2$ using the Cauchy-Schwarz inequality,
+\begin{align*}
+|L_{P,\varepsilon}u(x)|^2 & = \Biggl(\frac{1}{d_P(x)} \int_{\Omega} \bigl(u(x) - u(z)\bigr) \eta\biggl(\frac{|x - z|}{\varepsilon}\biggr) \,dz\Biggr)^2 \\
+& \leq \frac{1}{d_P(x)}\biggl(\int_{\Omega} K^{1/2}\biggl(\frac{x - z}{\varepsilon}\biggr) \,dz\biggr) \cdot \frac{1}{d_P(x)}\int_{\Omega} \bigl(u(x) - u(z)\bigr)^2 K^{1/2}\biggl(\frac{|x - z|}{\varepsilon}\biggr) \,dz \\
+& \leq 2\frac{\|\eta\|_{\infty}}{\alpha^2\varepsilon} \cdot \int_{\Omega} \bigl(u(x)^2 + u(z)^2\bigr) K^{1/2}\biggl(\frac{|x - z|}{\varepsilon}\biggr) \,dz,
+\end{align*}
+where in the final inequality we have used the estimates $d_P(x;K^{1/2}) \leq \|\eta\|_{\infty} \varepsilon$ and $d_P(x) \geq (\varepsilon\alpha)$.
+
+We use these pointwise estimates to upper bound~\eqref{pf:kernel_smoothing_bias_1d_2}.
+Since $\eta$ is compactly supported, if $x \in (0,\varepsilon]$ and $\eta(|x - z|/\varepsilon) > 0$ then $z \in (0,2\varepsilon]$.
+As a result,
+\begin{align*}
+\int_{0}^{\varepsilon} |L_{P,\varepsilon}u(x)|^2 \,dP(x) & \leq 2\frac{\|\eta\|_{\infty}}{\alpha^2\varepsilon} \int_{0}^{\varepsilon}  \int_{\Omega} \bigl(u(x)^2 + u(z)^2\bigr) K^{1/2}\biggl(\frac{|x - z|}{\varepsilon}\biggr) \,dz \,dx \\
+& \leq 2\frac{\|\eta\|_{\infty}}{\alpha^2\varepsilon} \int_{0}^{2\varepsilon}  \int_{0}^{2\varepsilon} \bigl(u(x)^2 + u(z)^2\bigr) K^{1/2}\biggl(\frac{|x - z|}{\varepsilon}\biggr) \,dz \,dx \\
+& \leq 4\frac{\|\eta\|_{\infty}^2}{\alpha^2} \int_{0}^{2\varepsilon} (u(x))^2 \,dx,
+\end{align*}
+where the last inequality follows by Fubini's Theorem.
+
+It remains to upper bound the $\Leb^2$ norm of $u$ over $(0,2\varepsilon]$. Here we use the bound~\eqref{pf:kernel_smoothing_bias_1d_3} and the Cauchy-Schwarz inequality to conclude that 
+\begin{equation*}
+\int_{0}^{2\varepsilon} \bigl(u(x)\bigr)^2 \,dx \leq \int_{0}^{2\varepsilon} \biggl(\int_{0}^{x} u^{(k)}(s)(x - s)^{k - 1} \,ds\biggr)^2 \,dx \leq \varepsilon^{2k - 1} \int_{0}^{2\varepsilon} \int_{0}^{x} \bigl(u^{(k)}(s)\bigr)^2 \,ds \,dx.
+\end{equation*}
+Finally, switching the order of integration and substituting $v = s/(2\varepsilon)$ gives
+\begin{align*}
+\int_{0}^{2\varepsilon} \int_{0}^{x} \bigl(u^{(k)}(s)\bigr)^2 \,ds \,dx = \int_{0}^{2\varepsilon} \int_{x}^{2\varepsilon} \bigl(u^{(k)}(s)\bigr)^2 \,dx \,ds \leq 2\varepsilon \int_{0}^{2\varepsilon} \bigl(u^{(k)}(s)\bigr)^2 \,ds = 4\varepsilon^2 \|u^{(k)}\|_P^2,
+\end{align*}
+establishing~\eqref{pf:kernel_smoothing_bias_1d_2} upon choosing $C = \frac{16\|\eta\|_{\infty}^2}{\alpha^2}$ in the statement of the inequality.
+
+\subsection{Proof of Lemma~\ref{lem:kernel_smoothing_bias}, $d = 1$ and $p \in C^k$.}
+\label{subsec:kernel_smoothing_bias_pf_1d_smooth_p}
+As in~\ref{subsec:kernel_smoothing_bias_pf_1d_uniform_p}, we take $\Omega = (0,1)$, let
+\begin{equation*}
+\|L_{P,\varepsilon}u\|_{P}^2 = \underbrace{\int_{\Omega_{\varepsilon}} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_1} + \underbrace{\int_{\partial_{\varepsilon}\Omega} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_2},
+\end{equation*}
+and turn to separately upper bounding $I_1$ and $I_2$.
+
+\paragraph{Upper bound on $I_1$.}
+We will show that for all $x \in \Omega_{\varepsilon}$,
+\begin{equation*}
+L_{P,\varepsilon}u(x) = \varepsilon^{k} \frac{\sigma_\eta}{k!}\Bigl( \frac{d^k}{dx^k}(up)(x) - p^{(k)}(x)u(x) + r(x)\Bigr)
+\end{equation*}
+for a remainder term $r$ satisfying $\int_{\Omega_{\varepsilon}} \bigl(r(x)\bigr)^2 p(x) \,dx  = o(1)$ as $\varepsilon \to 0$. This in turn will imply
+\begin{align*}
+I_1 = \int_{\Omega_{\varepsilon}} |L_{P,\varepsilon}u(x)|^2 \,dx & = \varepsilon^{2k} \frac{\sigma_\eta^2}{(k!)^2}\Bigl(\int_{\Omega_{\varepsilon}} \Bigl( \frac{d^k}{dx^k}(up)(x) - p^{(k)}(x)u(x) + r(x)\Bigr)^2  p(x) \,dx \\
+& \leq 3 \varepsilon^{2k} \frac{\sigma_\eta^2}{(k!)^2} \bigl( |up|_{H^k(P)}^2 + |p^{(k)}u|_{\Leb^2(P)}^2 + o(1)\bigr) \\
+& \leq C \varepsilon^{2k} \frac{\sigma_\eta^2}{(k!)^2} \Bigl( \|p\|_{C^k(P)}^2\bigl(|u|_{H^k(P)}^2 + |u|_{\Leb^2(P)}^2\bigr) + o(1) \Bigr),
+\end{align*}
+with the last line following from Young's inequality.
+
+It remains to show the pointwise estimate \textcolor{red}{(...)}.
 

### 2021-03-07 manuscripts/laplacian_smoothing/arXiv/appendix.tex
till just before B.2

@@
-\noindent In the appendix, we provide proofs of all results found in the paper ``Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs''.
-Our main theorems (Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing_manifold}) all follow the same general proof strategy, centered on conditioning.
+% !TEX root = minimax_optimal_laplacian_smoothing.tex
+\section{Preliminaries}
+\label{app:preliminaries}
+\noindent In the appendix, we provide complete proofs of all results. % found in the paper ``Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs''. 
+Our main theorems (Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing_manifold}) all follow the same general proof strategy of first establishing bounds in the fixed-design setup.
 In Section~\ref{sec:fixed_graph_error_bounds}, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to (functionals of) the graph $G$, and allow us to upper bound the error of $\wh{f}$ and $\wh{\varphi}$ conditional on the design $\{X_1,\ldots,X_n\} = \{x_1,\ldots,x_n\}$.
-In Sections~\ref{sec:graph_sobolev_seminorm}, \ref{sec:graph_eigenvalues}, \ref{sec:empirical_norm}, and \ref{sec:manifold} we give all our necessary probabilistic estimates on these functionals, for the particular random neighborhood graph $G = G_{n,r}$.
+In Sections~\ref{sec:graph_sobolev_seminorm}, \ref{sec:graph_eigenvalues}, \ref{sec:empirical_norm}, and \ref{sec:manifold} we develop all the necessary probabilistic estimates on these functionals, for the particular random neighborhood graph $G = G_{n,r}$.
 It is in these sections where we invoke our various assumptions on the distribution $P$ and regression function $f_0$.
@@
 
-\vfill
-
-\clearpage
-
 \section{Graph-dependent error bounds}
@@
 \label{eqn:ls_ts_G}
-\wh{T} := \frac{1}{n} \Bigl\|\wh{f}\Bigr\|_2^2.
+\wh{T} := \frac{1}{n} \|\wh{f}\|_2^2.
 \end{equation*}
@@
 
-Let $S \in \Reals^{n \times n}$ be a square, symmetric matrix, and let 
+Let $S \in \Reals^{n \times n}$ be a fixed square, symmetric matrix, and let 
 \begin{equation*}
@@
 	\begin{equation}
-	\wc{t}_{b} := \norm{\lambdavec(S)}_2^2 + 2b \norm{\lambdavec(S)}_4^2
+	\wc{t}_{b} := \norm{\lambdavec(S)}_2^2 + 2b \norm{\lambdavec(S)}_4^2.
 	\end{equation}
@@
 		\label{eqn:linear_smoother_fixed_graph_testing_typeI}
-		\Pbb_0\bigl(\wc{T} > \wc{t}_b\bigr) \leq \frac{1}{b^2}
+		\Pbb_0\bigl(\wc{T} > \wc{t}_b\bigr) \leq \frac{1}{b^2}.
 		\end{equation}
@@
 \frac{1}{n}\bigl\|\wc{f} - f_0\bigr\|_2^2 & \leq \frac{2}{n}\Bigl(\bigl\|\wc{f} - \Ebb_{f_0}[\wc{f}]\bigr\|_2^2 + \bigl\|\Ebb_{f_0}[\wc{f}] - f_0\bigr\|_2^2\Bigr) \\ 
-& = \frac{2}{n}\Bigl(\bigl\|S\varepsilon\bigr\|_2^2 + \bigl\|(S - I)f_0\bigr\|_2^2\Bigr)
-\end{align*}
+& = \frac{2}{n}\Bigl(\bigl\|S\varepsilon\bigr\|_2^2 + \bigl\|(S - I)f_0\bigr\|_2^2\Bigr).
+\end{align*}
+\sbcomment{General comment. Please punctuate math. I've done it till this point but please check everywhere.}
 Writing $\norm{S\varepsilon}_2^2 = \sum_{k = 1}^{n} \lambda_k(S)^2 Z_k^2$, the claim follows from the result of \citet{laurent00} on concentration of $\chi^2$-random variables, which for completeness we restate in Lemma~\ref{lem:chi_square_bound}.
@@
 \Pbb_{f_0}\Bigl(\wc{T} \leq \wc{t}_b\Bigr) & = \Pbb_{f_0}\Bigl(\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr] \leq \wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr) \\
-& \overset{(i)}{\leq} \Pbb_{f_0}\Bigl(\Bigl|\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr| \geq \Bigl|\wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr|\Bigr) \leq \frac{Var_{f_0}\bigl[\wc{T}\bigr]}{\bigl(\wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\bigr)^2} \\
+& \overset{(i)}{\leq} \Pbb_{f_0}\Bigl(\Bigl|\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr| \geq \Bigl|\wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr|\Bigr) \leq \frac{\Var_{f_0}\bigl[\wc{T}\bigr]}{\bigl(\wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\bigr)^2} \\
 & \overset{(ii)}{\leq} 4 \frac{\Var_{f_0}\bigl[\wc{T}\bigr]}{(f_0^{\top} S^2 f_0)^2} \\
@@
 \end{equation*}
+\sbcomment{I think the first equality should be the negative of what you wrote?}
 $(iii)$ follows from assumption \eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and the inequality \eqref{pf:linear_smoother_fixed_graph_testing_var}, and $(iv)$ follows from \eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}.

### 2021-03-07 manuscripts/laplacian_smoothing/arXiv/appendix.tex
Changed proof of Lemma 4 in response to Siva's comment.

@@
 In Lemma~\ref{lem:linear_smoother_fixed_graph_testing}, we upper bound the error of a test involving the statistic $\|\wc{f}\|_2^2 = {\bf Y}^{\top} S^2 {\bf Y}$.
+We will require that $S$ be a \emph{contraction}, meaning that it has operator norm no greater than $1$, $\|Sv\|_2 \leq \|v\|_2$ for all $v \in \Reals^n$.
 \begin{lemma}
 	\label{lem:linear_smoother_fixed_graph_testing}
-	Let $\wc{T} = Y^{\top} S^2 Y$ for a square, symmetric matrix $S \in \Reals^{n \times n}$. Define the threshold $\wc{t}_b$ to be 
-	\begin{equation}
+	Let $\wc{T} = {\bf Y}^{\top} S^2 {\bf Y}$ for a square, symmetric matrix $S \in \Reals^{n \times n}$. Suppose $S$ is a contraction. Define the threshold $\wc{t}_b$ to be 
+	\begin{equation}
+	\label{eqn:linear_smoother_fixed_graph_threshold}
 	\wc{t}_{b} := \norm{\lambdavec(S)}_2^2 + 2b \norm{\lambdavec(S)}_4^2.
@@
 
-\textit{Variance.} Starting from \eqref{pf:linear_smoother_fixed_graph_testing1} and recalling $\Var(Z_k^2) = 2$, we derive
+\textit{Variance.} We start from \eqref{pf:linear_smoother_fixed_graph_testing1}. Recalling that $\Var(Z_k^2) = 2$, it follows from the Cauchy-Schwarz inequality that
 \begin{equation}
@@
 
-\textit{Bounding Type I and Type II error.} The upper bound \eqref{eqn:linear_smoother_fixed_graph_testing_typeI} follows immediately from \eqref{pf:linear_smoother_fixed_graph_testing_mean}, \eqref{pf:linear_smoother_fixed_graph_testing_var}, and Chebyshev's inequality.
-
-The bound \eqref{eqn:linear_smoother_fixed_graph_testing_typeII} also follows from Chebyshev's inequality, as can be seen by the following manipulations,
-\begin{equation*}
-\begin{aligned}
+\textit{Bounding Type I and Type II error.} The upper bound \eqref{eqn:linear_smoother_fixed_graph_testing_typeI} on Type I error follows immediately from \eqref{pf:linear_smoother_fixed_graph_testing_mean}, \eqref{pf:linear_smoother_fixed_graph_testing_var}, and Chebyshev's inequality.
+
+We now establish the upper bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeII} on Type II error.
+From assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}, we see that $f_0^{\top} S^2 f_0^{\top} - \wc{t}_b \leq 0$.
+As a result,
+\begin{align*}
 \Pbb_{f_0}\Bigl(\wc{T} \leq \wc{t}_b\Bigr) & = \Pbb_{f_0}\Bigl(\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr] \leq \wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr) \\ 
-& \overset{(i)}{\leq} \Pbb_{f_0}\Bigl(\Bigl|\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr| \geq \Bigl|\wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr|\Bigr) \leq \frac{\Var_{f_0}\bigl[\wc{T}\bigr]}{\bigl(\wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\bigr)^2} \\
-& \overset{(ii)}{\leq} 4 \frac{\Var_{f_0}\bigl[\wc{T}\bigr]}{(f_0^{\top} S^2 f_0)^2} \\
-& \overset{(iii)}{\leq} \frac{32}{f_0^{\top} S^2 f_0} + \frac{4}{b^2} \\
-& \overset{(iv)}{\leq} \frac{8}{b} \Biggl(\sum_{k = 1}^{n}\bigl(\lambda_k(S)\bigr)^4\Biggr)^{-1/2} + \frac{4}{b^2}
-\end{aligned}
-\end{equation*}
-In the previous expression, $(i)$ and $(ii)$ follow since assumption \eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and equation \eqref{pf:linear_smoother_fixed_graph_testing_mean} together imply
-\begin{equation*}
-\wc{t}_b - \Ebb_{f_0}[\wc{T}] = f_0^{\top} S^2 f_0 - 2b\|{\bf \lambda}(S)\|_4^2 \geq \frac{1}{2}f_0^{\top} S^2 f_0
-\end{equation*}
+& \leq \Pbb_{f_0}\Bigl(\Bigl|\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr| \geq \Bigl|\wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr|\Bigr) \\ 
+& \leq \frac{\Var_{f_0}\bigl[\wc{T}\bigr]}{\bigl(\wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\bigr)^2},
+\end{align*}
+where the last line follows from Chebyshev's inequality.
+Plugging in the expressions~\eqref{pf:linear_smoother_fixed_graph_testing_mean} and~\eqref{pf:linear_smoother_fixed_graph_testing_var} for the mean and variance of $\wc{T}$, as well as the definition of $\wc{t}_b$ in~\eqref{eqn:linear_smoother_fixed_graph_threshold}, we obtain that
+\begin{equation}
+\label{pf:linear_smoother_fixed_graph_testing2}
+\Pbb_{f_0}\Bigl(\wc{T} \leq \wc{t}_b\Bigr) \leq \frac{4\|\lambdavec(S)\|_4^4}{\bigl(f_0^{\top}S^2f_0 - 2b \norm{\lambdavec(S)}_4^2\bigr)^2} + \frac{8f_0^{\top}S^4f_0 }{\bigl(f_0^{\top}S^2f_0 - 2b \norm{\lambdavec(S)}_4^2\bigr)^2}.
+\end{equation}
+We now use the assumed lower bound $f_0^{\top} S^2 f_0 \geq 4b \norm{\lambdavec(S)}_4^2$ to separately upper bound each of the two terms on the right hand side of~\eqref{pf:linear_smoother_fixed_graph_testing2}.
+It follows immediately that
+\begin{equation}
+\label{pf:linear_smoother_fixed_graph_testing2.5}
+\frac{4 \|\lambdavec(S)\|_4^4}{\bigl(f_0^{\top}S^2f_0 - 2b \norm{\lambdavec(S)}_4^2\bigr)^2} \leq \frac{1}{b^2},
+\end{equation}
+giving a sufficient upper bound on the first term. Now we upper bound the second term,
+\begin{equation}
+\label{pf:linear_smoother_fixed_graph_testing3}
+\frac{8f_0^{\top}S^4f_0 }{\bigl(f_0^{\top}S^2f_0 - 2b \norm{\lambdavec(S)}_4^2\bigr)^2} \leq \frac{32 f_0^{\top}S^4f_0 }{\bigl(f_0^{\top}S^2f_0\bigr)^2} \leq \frac{8}{b \|\lambdavec(S)\|_4^2} \frac{f_0^{\top}S^4f_0 }{f_0^{\top}S^2f_0} \leq \frac{8}{b \|\lambdavec(S)\|_4^2},
+\end{equation}
+where the final inequality is satisfied because $S$ is a contraction.
+Plugging~\eqref{pf:linear_smoother_fixed_graph_testing2.5} and~\eqref{pf:linear_smoother_fixed_graph_testing3} back into~\eqref{pf:linear_smoother_fixed_graph_testing2} then gives the desired result.
+
 \sbcomment{I think the first equality should be the negative of what you wrote?}
-$(iii)$ follows from assumption \eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and the inequality \eqref{pf:linear_smoother_fixed_graph_testing_var}, and $(iv)$ follows from \eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}.
-
+
+\agcomment{You were correct.
+However I also rewrote the proof a bit to make it (hopefully) easier to follow.
+We also need that $S$ is a contraction---a condition that I had and then deleted at some point by accident--- and I added this to the statement of the Lemma.}
 \subsection{Analysis of Laplacian smoothing}
@@
 	\label{eqn:ls_fixed_graph_estimation_prob}
-	\frac{1}{n}\bigl\|\wh{f} - f_0\bigr\|_2^2 \leq \frac{2\rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{10}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k} + 1\bigr)^2}
+	\frac{1}{n}\bigl\|\wh{f} - f_0\bigr\|_2^2 \leq \frac{2\rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{10}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_{k} + 1\bigr)^2},
 	\end{equation}
@@
 		\label{eqn:ls_fixed_graph_testing_typeI}
-		\Pbb_0\Bigl(\wh{T} > \wh{t}_b\Bigr) \leq \frac{1}{b^2}
+		\Pbb_0\Bigl(\wh{T} > \wh{t}_b\Bigr) \leq \frac{1}{b^2}.
 		\end{equation}
@@
 		\label{eqn:ls_fixed_graph_testing_critical_radius}
-		\frac{1}{n}\norm{f_0}_2^2 \geq \frac{2 \rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{4b}{n} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{1/2}
+		\frac{1}{n}\norm{f_0}_2^2 \geq \frac{2 \rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{4b}{n} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{1/2},
 		\end{equation}
@@
 		\label{eqn:ls_fixed_graph_testing_typeII}
-		\Pbb_{f_0}\Bigl(\wh{T}(G) \leq \wh{t}_b\Bigr) \leq \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2}
+		\Pbb_{f_0}\Bigl(\wh{T}(G) \leq \wh{t}_b\Bigr) \leq \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2}.
 		\end{equation}
@@
 \begin{aligned}
-\bigl\|(\wh{S} - I)f_0\bigr\|_2^2 & = f_0^{\top} \Lap^{1/2} \Lap^{-1/2}\bigl(\wh{S} - \Id\bigr) \Lap^{-1/2} \Lap^{1/2} f_0 \\
-& \leq f_0^{\top} \Lap f_0 \cdot \lambda_n\Bigl(\Lap^{-1/2}\bigl(\wh{S} - \Id\bigr)\Lap^{-1/2}\Bigr) \\
-& = f_0^{\top} \Lap f_0 \cdot \max_{k \in [n]} \biggl\{\frac{1}{\lambda_k} \Bigl(1 - \frac{1}{\rho\lambda_k + 1}\Bigr) \biggr\} \\
-& = f_0^{\top} \Lap f_0 \cdot \rho.
+\bigl\|(\wh{S} - I)f_0\bigr\|_2^2 & = f_0^{\top} \Lap^{1/2} \Lap^{-1/2}\bigl(\wh{S} - \Id\bigr)^2 \Lap^{-1/2} \Lap^{1/2} f_0 \\
+& \leq f_0^{\top} \Lap f_0 \cdot \lambda_n\Bigl(\Lap^{-1/2}\bigl(\wh{S} - \Id\bigr)^2 \Lap^{-1/2}\Bigr) \\
+& = f_0^{\top} \Lap f_0 \cdot \max_{k \in [n]} \biggl\{\frac{1}{\lambda_k} \Bigl(1 - \frac{1}{\rho\lambda_k + 1}\Bigr)^2 \biggr\} \\
+& \leq f_0^{\top} \Lap f_0 \cdot \rho.
 \end{aligned}
@@
 \paragraph{Proof of Lemma~\ref{lem:ls_fixed_graph_testing}.}
-Let $\wh{S} := (\rho \Lap + \Id)^{-1}$, and note that $\wh{T} = \frac{1}{n}{\bf Y}^{\top} \wh{S}^2 {\bf Y}$.
+Let $\wh{S} := (I + \rho L)^{-1}$, so that $\wh{T} = \frac{1}{n}{\bf Y}^{\top} \wh{S}^2 {\bf Y}$.
+Note that $\wh{S}$ is a contraction, so that we may invoke Lemma~\ref{lem:linear_smoother_fixed_graph_testing}.
 The bound on Type I error \eqref{eqn:ls_fixed_graph_testing_typeI} follows immediately from \eqref{eqn:linear_smoother_fixed_graph_testing_typeI}.
@@
 & \leq f_0^{\top} \Lap f_0 \cdot  \lambda_{n}\biggl(\Lap^{-1/2}\Bigl(\Id - \wh{S}^2\Bigr) \Lap^{-1/2}\biggr) \\ 
-& \overset{(i)}{=}  f_0^{\top} \Lap f_0 \cdot \max_{k} \biggl\{ \frac{1}{\lambda_k} \Bigl(1 - \frac{1}{(\rho \lambda_k + 1)^2}\Bigr) \biggr\} \\
-& \overset{(ii)}{\leq} f_0^{\top} \Lap f_0 \cdot 2\rho,
+& = f_0^{\top} \Lap f_0 \cdot \max_{k} \biggl\{ \frac{1}{\lambda_k} \Bigl(1 - \frac{1}{(\rho \lambda_k + 1)^2}\Bigr) \biggr\} \\
+& \leq f_0^{\top} \Lap f_0 \cdot 2\rho,
 \end{aligned}
@@
 \begin{equation*}
-f_0^{\top} \wh{S}^2 f_0 \geq 2b \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{-1/2}.
+f_0^{\top} \wh{S}^2 f_0 \geq 4b \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{-1/2}.
 \end{equation*} 
@@
 \label{pf:first_order_graph_sobolev_seminorm_1}
-\Ebb\Bigl[f^{\top} \Lap_{n,r} f\Bigr] = \frac{n(n - 1)}{2} \Ebb\biggl[\Bigl(f(X') - f(X)\Bigr)^2 K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr]
+\Ebb\Bigl[f^{\top} \Lap_{n,r} f\Bigr] = \frac{n(n - 1)}{2} \Ebb\biggl[\Bigl(f(X') - f(X)\Bigr)^2 K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr],
 \end{equation}
@@
 & \overset{(ii)}{\leq} r^{d + 2} \int_{\Xset} \int_{B(0,1)} \int_{0}^{1} \biggl[\nabla f\bigl(x + trz\bigr)^{\top} z\biggr]^2 K\bigl(\|z\|\bigr) \,dt\,dz \,dx \nonumber \\
-&  \overset{(iii)}{\leq} r^{d + 2} \int_{\Omega} \int_{B(0,1)} \int_{0}^{1} \Bigl[\nabla f\bigl(\wt{x}\bigr)^{\top} z\Bigr]^2 K\bigl(\|z\|\bigr) \,dt \,dz \,d\wt{x} \label{pf:first_order_graph_sobolev_seminorm_3}
+&  \overset{(iii)}{\leq} r^{d + 2} \int_{\Omega} \int_{B(0,1)} \int_{0}^{1} \Bigl[\nabla f\bigl(\wt{x}\bigr)^{\top} z\Bigr]^2 K\bigl(\|z\|\bigr) \,dt \,dz \,d\wt{x}, \label{pf:first_order_graph_sobolev_seminorm_3}
 \end{align}
@@
 	\begin{equation*}
-	f^{\top} \Lap_{n,r} f \leq \biggl(\frac{1}{\delta} + C_2\biggr) n^2 r^{d + 2} M^2
+	f^{\top} \Lap_{n,r} f \leq \biggl(\frac{1}{\delta} + C_2\biggr) n^2 r^{d + 2} M^2.
 	\end{equation*}

### 2021-03-07 manuscripts/laplacian_smoothing/arXiv/appendix.tex
Added punctuation to math in appendix.

@@
 	& \leq p_{\max}^2 M^4 r^4 K(0) \int_{\Xset} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx' \,dx \\
-	& \leq p_{\max}^2 M^4 r^{4 + d} K(0)
+	& \leq p_{\max}^2 M^4 r^{4 + d} K(0).
 	\end{align*}
@@
 \begin{equation*}
-\Pbb\biggl(\Bigl|f^{\top} \Lap_{n,r} f - \Ebb\bigl[f^{\top} \Lap_{n,r} f\bigr]\Bigr| \geq \frac{M^2}{\delta}n^2 r^{d + 2}\biggr) \leq C\frac{\delta^2}{n}
+\Pbb\biggl(\Bigl|f^{\top} \Lap_{n,r} f - \Ebb\bigl[f^{\top} \Lap_{n,r} f\bigr]\Bigr| \geq \frac{M^2}{\delta}n^2 r^{d + 2}\biggr) \leq C\frac{\delta^2}{n},
 \end{equation*}
@@
 \begin{equation*}
-\Delta_Pf(x) := -\frac{1}{p(x)} \dive\bigl(p^2\nabla f\bigr)(x)
+\Delta_Pf(x) := -\frac{1}{p(x)} \dive\bigl(p^2\nabla f\bigr)(x).
 \end{equation*}
@@
 We let $L_K$ denote the Lipschitz constant of the kernel $K$.
-Finally, for notational ease we set $\theta$ and $\wt{\delta}$ to be the following (small) positive constants.
+Finally, for notational ease we set $\theta$ and $\wt{\delta}$ to be the following (small) positive numbers:
 \begin{equation}
@@
 
-It is therefore the case that the inequalities in \eqref{eqn:eigenvalue_bound} hold with probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{m})$.
+It is therefore the case that the inequalities in \eqref{eqn:eigenvalue_bound} hold with probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{d})$.
 Together, \eqref{eqn:eigenvalue_bound} and \eqref{eqn:weyls_law} imply the following bounds on the graph Laplacian eigenvalues:
@@
 \begin{equation*}
-\lambda_k(G_{n,r}) \leq 2D_{\max}(G_{n,r}) \leq 4p_{\max}nr^d 
+\lambda_k(G_{n,r}) \leq 2D_{\max}(G_{n,r}) \leq 4p_{\max}nr^d,
 \end{equation*}
@@
 	\label{prop:dirichlet_energies}
-	With probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{m})$, we have the following.
+	With probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{d})$, we have the following.
 	\begin{enumerate}[(1)]
@@
 		\label{eqn:dirichlet_energies_1}
-		\sigma_{K} D_2(\wt{\mc{I}}u) \leq A_8 \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) b_r(u)
+		\sigma_{K} D_2(\wt{\mc{I}}u) \leq A_8 \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) b_r(u).
 		\end{equation}
@@
 		\label{eqn:dirichlet_energies_2}
-		b_r(\wt{\mc{P}}f) \leq \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr) \cdot \Bigl(\frac{C_5 p_{\max}^2}{p_{\min}^2}\Bigr) \cdot \sigma_{K} D_2(f)
+		b_r(\wt{\mc{P}}f) \leq \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr) \cdot \Bigl(\frac{C_5 p_{\max}^2}{p_{\min}^2}\Bigr) \cdot \sigma_{K} D_2(f).
 		\end{equation}
@@
 	\label{prop:isometry}
-	With probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{m})$, we have the following.
+	With probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{d})$, we have the following.
 	\begin{enumerate}[(1)]
@@
 		\label{eqn:isometry_1}
-		\Bigl|\|f\|_{P}^2 - \|\wt{\mc{P}}f\|_{n}^2\Bigr| \leq A_5 r \|f\|_{P} \sqrt{D_2(f)} + A_1\bigl(\theta + \wt{\delta}\bigr)\|f\|_{P}^2
+		\Bigl|\|f\|_{P}^2 - \|\wt{\mc{P}}f\|_{n}^2\Bigr| \leq A_5 r \|f\|_{P} \sqrt{D_2(f)} + A_1\bigl(\theta + \wt{\delta}\bigr)\|f\|_{P}^2.
 		\end{equation}
@@
 		\label{eqn:isometry_2}
-		\Bigl|\|\wt{\mc{I}}u\|_{P}^2 - \|u\|_{n}^2\Bigr| \leq A_6 r \|u\|_{n} \sqrt{b_r(u)} + A_7\bigl(\theta + \wt{\delta}\bigr) \|u\|_{n}^2
+		\Bigl|\|\wt{\mc{I}}u\|_{P}^2 - \|u\|_{n}^2\Bigr| \leq A_6 r \|u\|_{n} \sqrt{b_r(u)} + A_7\bigl(\theta + \wt{\delta}\bigr) \|u\|_{n}^2.
 		\end{equation}
@@
 & \geq \|u\|_n^2 - A_6 r \|u\|_n^2 \sqrt{\frac{1}{a}\lambda_k(\Delta_P)} - A_7\bigl(\theta + \wt{\delta}\bigr)\|u\|_n^2 \\
-& \geq \frac{1}{2}\|u\|_n^2
+& \geq \frac{1}{2}\|u\|_n^2,
 \end{align*}
@@
 & \leq 2A_8\biggl(1 + A_1(\theta + \wt{\delta})\biggr) \biggl(1 + A_3 \frac{\wt{\delta}}{r}\biggr)\max_{u \in U} \frac{b_r(u)}{\|u\|_n^2} \\
-& \leq 2A_8\biggl(1 + A_1(\theta + \wt{\delta})\biggr) \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) \frac{\lambda_k(G_{n,r})}{nr^{d + 2}}
+& \leq 2A_8\biggl(1 + A_1(\theta + \wt{\delta})\biggr) \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) \frac{\lambda_k(G_{n,r})}{nr^{d + 2}},
 \end{align*}
@@
 \begin{equation*}
-{E}_r(f,\eta,V) := \frac{1}{r^{d + 2}}\int_{V} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx
-\end{equation*}
-where $\eta: [0,\infty) \to [0,\infty)$ is an arbitrary kernel, and $V \subseteq \Xset$ is a measurable set.
+{E}_r(f,\eta,V) := \frac{1}{r^{d + 2}}\int_{V} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx.
+\end{equation*}
+Here $\eta: [0,\infty) \to [0,\infty)$ is an arbitrary kernel, and $V \subseteq \Xset$ is a measurable set.
 We will abbreviate $\wt{E}_r(f,\eta,\Xset)$ as $\wt{E}_r(f,\eta)$ and $\wt{E}_r(f,K) = \wt{E}_r(f)$ (and likewise with $E_r$.)
@@
 \label{pf:prop_optimal_transport_0}
-\wt{p}_n(x) := \frac{P_n(Q)}{\vol(Q)}
+\wt{p}_n(x) := \frac{P_n(Q)}{\vol(Q)},
 \end{equation}
@@
 \begin{itemize}
-	\item In the special case of the unit cube $\Xset = (0,1)^d$, the partition will simply be the collection of cubes
-	\begin{equation*}
-	\set{Q_k: k \in [\wt{\delta}^{-1}]^d} 
+	\item In the special case of the unit cube $\Xset = (0,1)^d$, the partition will simply be a collection of cubes,
+	\begin{equation*}
+	\mc{Q} = \set{Q_k: k \in [\wt{\delta}^{-1}]^d},
 	\end{equation*}
@@
 Moreover, letting $U_j$ denote the intersection of the $j$th of these polytopes with $\wb{\Xset}$, this proposition establishes that for each $j$ there exists a bi-Lipschitz homeomorphism $\Phi_j: U_j \to [0,1]^d$.
-We take the collection of
+We take the collection
 	\begin{equation*}
@@
 \label{pf:prop_optimal_transport_1}
-\diam(\wt{Q}_{j,k}) \leq L_{\Psi} \L_{\Phi} \wt{\delta}
+\diam(\wt{Q}_{j,k}) \leq L_{\Psi} \L_{\Phi} \wt{\delta},
 \end{equation}
@@
 \begin{align*}
-|\wt{p}_n(x) - p(x)| \leq \biggl|\frac{P_n(\wt{Q}_{j,k}) - P(\wt{Q}_{j,k})}{\vol(\wt{Q}_{j,k})}\biggr| + \frac{1}{\vol(\wt{Q}_{j,k})} \int_{\wt{Q}_{j,k}} |p(x') - p(x)| \,dx 
+|\wt{p}_n(x) - p(x)| \leq \biggl|\frac{P_n(\wt{Q}_{j,k}) - P(\wt{Q}_{j,k})}{\vol(\wt{Q}_{j,k})}\biggr| + \frac{1}{\vol(\wt{Q}_{j,k})} \int_{\wt{Q}_{j,k}} |p(x') - p(x)| \,dx,
 \end{align*}
@@
 \label{pf:prop_optimal_transport_3}
-\|\wt{p}_n - p\|_{\Leb^{\infty}} \leq \max_{j,k} \biggl|\frac{P_n(\wt{Q}_{j,k}) - P(\wt{Q}_{j,k})}{\vol(\wt{Q}_{j,k})}\biggr| + L_p L_{\Phi} L_{\Psi} \wt{\delta}
+\|\wt{p}_n - p\|_{\Leb^{\infty}} \leq \max_{j,k} \biggl|\frac{P_n(\wt{Q}_{j,k}) - P(\wt{Q}_{j,k})}{\vol(\wt{Q}_{j,k})}\biggr| + L_p L_{\Phi} L_{\Psi} \wt{\delta}.
 \end{equation}
@@
 	\label{eqn:smoothening_error_norm}
-	\bigl\|\Lambda_hf\bigr\|_{P}^2 \leq \frac{p_{\max}}{a_3p_{\min}} \bigl\|f\bigr\|_{P}^2
+	\bigl\|\Lambda_hf\bigr\|_{P}^2 \leq \frac{p_{\max}}{a_3p_{\min}} \bigl\|f\bigr\|_{P}^2,
 	\end{equation}
@@
 	\label{eqn:smoothening_error_energy}
-	\bigl\|\Lambda_hf - f\bigr\|_{P}^2 \leq \frac{1}{a_3\sigma_Kp_{\min}} h^2 E_h(f)
+	\bigl\|\Lambda_hf - f\bigr\|_{P}^2 \leq \frac{1}{a_3\sigma_Kp_{\min}} h^2 E_h(f),
 	\end{equation}
@@
 \begin{equation*}
-\Lambda_hf(x) = a + \frac{1}{h^d\tau_h(x)}\int_{\Xset} \eta_h(x',x)\bigl(f(x') - a\bigr)\,dx'
+\Lambda_hf(x) = a + \frac{1}{h^d\tau_h(x)}\int_{\Xset} \eta_h(x',x)\bigl(f(x') - a\bigr)\,dx',
 \end{equation*}
@@
 \begin{equation*}
-\|\nabla\tau_h(x)\|^2 \leq \frac{1}{\sigma_K^2 h^{2d + 4}} \biggl[\int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr)\,dx'\biggr] \biggl[\int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr)\|x' - x\|^2\,dx',\biggr] \leq \frac{1}{d\sigma_K h^{2}}
+\|\nabla\tau_h(x)\|^2 \leq \frac{1}{\sigma_K^2 h^{2d + 4}} \biggl[\int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr)\,dx'\biggr] \biggl[\int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr)\|x' - x\|^2\,dx',\biggr] \leq \frac{1}{d\sigma_K h^{2}},
 \end{equation*}
@@
 \begin{equation*}
-\bigl|\Lambda_rf(x) - f(x)\bigr|^2 \leq \frac{1}{h^{d} a_3 \sigma_K p_{\min}} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr) \bigl(f(x') - f(x)\bigr)^2 p(x') \,dx'
+\bigl|\Lambda_rf(x) - f(x)\bigr|^2 \leq \frac{1}{h^{d} a_3 \sigma_K p_{\min}} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{h}\biggr) \bigl(f(x') - f(x)\bigr)^2 p(x') \,dx'.
 \end{equation*}
@@
 	\begin{equation*}
-	b_r(\wt{\mc{P}}f) \leq \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr) \wt{E}_{r + 2A_0\wt{\delta}}(f)
+	b_r(\wt{\mc{P}}f) \leq \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr) \wt{E}_{r + 2A_0\wt{\delta}}(f).
 	\end{equation*}
@@
 	\begin{equation*}
-	\wt{E}_{r - 2A_0\wt{\delta}}\bigl(\wt{\mc{P}}^{\star}u\bigr) \leq \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) b_{r}(u)
+	\wt{E}_{r - 2A_0\wt{\delta}}\bigl(\wt{\mc{P}}^{\star}u\bigr) \leq \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) b_{r}(u).
 	\end{equation*}
@@
 & \leq \frac{2^{d + 1}}{a_3 \nu_d r^d} \int_{V \cap \Xset} \bigl|f(x') - f(z)\big|^2 + \bigl|f(z) - f(x)\big|^2 \,dz \nonumber \\
-& \leq \frac{2^{d + 2}}{K(1) a_3 \nu_d r^d p_{\min}} \Bigl(F(x') + F(x)\bigr) \label{pf:poincare_2}
+& \leq \frac{2^{d + 2}}{K(1) a_3 \nu_d r^d p_{\min}} \Bigl(F(x') + F(x)\bigr), \label{pf:poincare_2}
 \end{align}
@@
 \begin{equation*}
-K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) \leq K\biggl(\frac{\bigl(\|x' - x\| - 2A_0\wt{\delta}\bigr)_{+}}{r}\biggr) \leq K\biggl(\frac{\|x' - x\|}{r + 2A_0\wt{\delta}}\biggr) + \frac{2L_KA_0\wt{\delta}}{r}\1\Bigl\{\|x' - x\| \leq r + 2A_0\wt{\delta}\Bigr\}
+K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) \leq K\biggl(\frac{\bigl(\|x' - x\| - 2A_0\wt{\delta}\bigr)_{+}}{r}\biggr) \leq K\biggl(\frac{\|x' - x\|}{r + 2A_0\wt{\delta}}\biggr) + \frac{2L_KA_0\wt{\delta}}{r}\1\Bigl\{\|x' - x\| \leq r + 2A_0\wt{\delta}\Bigr\}.
 \end{equation*}
@@
 & \leq \frac{1}{r^{d + 2}} \sum_{i,j = 1}^n \int_{U_i} \int_{U_j}  \bigl(f(x') - f(x)\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) \biggl[K\biggl(\frac{\|x' - x\|}{r + 2A_0\wt{\delta}}\biggr) + \frac{2L_KA_0\wt{\delta}}{r}\1\Bigl\{\|x' - x\| \leq r + 2\wt{\delta}\Bigr\}\biggr] \,dx' \,dx \\
-& = \Bigl(1 + 2A_0\frac{\wt{\delta}}{r}\Bigr)^{d + 2}\biggl[\wt{E}_{r + 2A_0\wt{\delta}}(f) + \frac{2L_KA_0\wt{\delta}}{r}\wt{E}_{r + 2A_0\wt{\delta}}(f; \1_{[0,1]})\biggr]
+& = \Bigl(1 + 2A_0\frac{\wt{\delta}}{r}\Bigr)^{d + 2}\biggl[\wt{E}_{r + 2A_0\wt{\delta}}(f) + \frac{2L_KA_0\wt{\delta}}{r}\wt{E}_{r + 2A_0\wt{\delta}}(f; \1_{[0,1]})\biggr],
 \end{align*}
@@
 \begin{align*}
-\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & = \frac{1}{\wt{r}^{d + 2}} \sum_{i = 1}^{n} \sum_{j = 1}^{n} \int_{U_i} \int_{U_j} \Bigl(u(X_i) - u(X_j)\Bigr)^2 K\biggl(\frac{\|x' - x\|}{\wt{r}}\biggr) \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx
+\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & = \frac{1}{\wt{r}^{d + 2}} \sum_{i = 1}^{n} \sum_{j = 1}^{n} \int_{U_i} \int_{U_j} \Bigl(u(X_i) - u(X_j)\Bigr)^2 K\biggl(\frac{\|x' - x\|}{\wt{r}}\biggr) \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx.
 \end{align*}
@@
 \begin{equation*}
-K\biggl(\frac{\|x' - x\|}{\wt{r}}\biggr) \leq K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) + \frac{2A_0L_K\wt{\delta}}{\wt{r}} \cdot \1\biggl\{\frac{\|x' - x\|}{\wt{r}} \leq 1\biggr\}
+K\biggl(\frac{\|x' - x\|}{\wt{r}}\biggr) \leq K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) + \frac{2A_0L_K\wt{\delta}}{\wt{r}} \cdot \1\biggl\{\frac{\|x' - x\|}{\wt{r}} \leq 1\biggr\},
 \end{equation*}
@@
 & \leq \biggl(1 + 2^{d + 2}A_0\frac{\wt{\delta}}{r}\biggr) b_r(u) + \frac{2A_0L_K\wt{\delta}}{\wt{r}}\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \nonumber \\
-& \leq \biggl(1 + 2^{d + 2}A_0\frac{\wt{\delta}}{r}\biggr) b_r(u) + \frac{4A_0L_K\wt{\delta}}{K(1)r} \wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u\bigr) \nonumber 
+& \leq \biggl(1 + 2^{d + 2}A_0\frac{\wt{\delta}}{r}\biggr) b_r(u) + \frac{4A_0L_K\wt{\delta}}{K(1)r} \wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u\bigr). \nonumber 
 \end{align}
-or in other words
+In other words,
 \begin{align*}
 \wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & \leq \biggl(1 - \frac{4A_0L_K\wt{\delta}}{K(1)r}\biggr)^{-1}\biggl(1 + 2^{d + 2}A_0\frac{\wt{\delta}}{r}\biggr) b_r(u) \\
-& \leq \biggl(1 + \frac{\wt{\delta}}{r}\Bigl(\frac{8A_0L_K}{K(1)} + 2^{d + 3}\Bigr)\biggr) b_r(u)
+& \leq \biggl(1 + \frac{\wt{\delta}}{r}\Bigl(\frac{8A_0L_K}{K(1)} + 2^{d + 3}\Bigr)\biggr) b_r(u),
 \end{align*}
@@
 & \overset{(ii)}{\leq} A_8 \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \wt{E}_{r - 2A_0\wt{\delta}}(\wt{\mc{P}}^{\star}u) \\
-& \overset{(iii)}{\leq} A_8 \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) b_r(u)
+& \overset{(iii)}{\leq} A_8 \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \biggl(1 + A_3\frac{\wt{\delta}}{r}\biggr) b_r(u),
 \end{align*}
@@
 & \overset{(v)}{\leq} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr){E}_{r + 2A_0\wt{\delta}}(f) \\
-& \overset{(vi)}{\leq} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr) \cdot \Bigl(\frac{C_5 p_{\max}^2}{p_{\min}^2}\Bigr) \cdot \sigma_{K} D_2(f)
+& \overset{(vi)}{\leq} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + A_9\frac{\wt{\delta}}{r}\Bigr) \cdot \Bigl(\frac{C_5 p_{\max}^2}{p_{\min}^2}\Bigr) \cdot \sigma_{K} D_2(f),
 \end{align*}
@@
 & \leq n \int_{U_i} \bigl|f(x)\bigr|^2 \wt{p}_n(x) \,dx \\
-& \leq n \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)\biggl[\int_{U_i} \bigl|f(x)\bigr|^2 p(x) \,dx + A_1(\theta + \wt{\delta}) \int_{U_i} \bigl|f(x)\bigr|^2 p(x) \,dx\biggr]
+& \leq n \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)\biggl[\int_{U_i} \bigl|f(x)\bigr|^2 p(x) \,dx + A_1(\theta + \wt{\delta}) \int_{U_i} \bigl|f(x)\bigr|^2 p(x) \,dx\biggr],
 \end{align*}
@@
 & \overset{(ii)}{\leq} 2 \sqrt{1 + A_1(\theta + \wt{\delta})} \Bigl|\bigl\|\wt{P}f\bigr\|_{n} - \bigl\|f\bigr\|_{\wt{P}_n}\Bigr| \cdot \bigl\|f\bigr\|_{P} + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2 \nonumber \\
-& \leq 2 \sqrt{1 + A_1(\theta + \wt{\delta})} \bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\wt{P}_n} \cdot \bigl\|f\bigr\|_{P} + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2 \label{pf:prop_isometry_2}
+& \leq 2 \sqrt{1 + A_1(\theta + \wt{\delta})} \bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\wt{P}_n} \cdot \bigl\|f\bigr\|_{P} + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2, \label{pf:prop_isometry_2}
 \end{align}
@@
 \end{equation*}
-From Lemma~\ref{lem:poincare}, we have that
-\begin{equation*}
-\int_{U_i} \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx \leq A_3r^2 \wt{E}_r(f,U_i)
-\end{equation*}
-for each $i = 1,\ldots,n$. Summing up over $i$ on both sides of the inequality
-\begin{equation*}
-\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\wt{P}_n}^2 \leq A_3r^2 \wt{E}_r(f,\Xset) \leq  A_3 \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(\frac{C_5 p_{\max}^2}{p_{\min}^2}\Bigr) \cdot \sigma_{K}r^2 D_2(f)
+From Lemma~\ref{lem:poincare}, we have that for each $i = 1,\ldots,n$,
+\begin{equation*}
+\int_{U_i} \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx \leq A_3r^2 \wt{E}_r(f,U_i).
+\end{equation*}
+Summing up over $i$ on both sides of the inequality gives
+\begin{equation*}
+\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\wt{P}_n}^2 \leq A_3r^2 \wt{E}_r(f,\Xset) \leq  A_3 \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(\frac{C_5 p_{\max}^2}{p_{\min}^2}\Bigr) \cdot \sigma_{K}r^2 D_2(f),
 \end{equation*}
@@
 & \overset{(iv)}{\leq} \wt{r} \sqrt{\frac{1}{a_3\sigma_K p_{\min}} E_{\wt{r}}(\wt{\mc{P}}^{\star}u)} \nonumber \\
-& \overset{(v)}{\leq} \wt{r} \sqrt{\frac{1 + A_1(\theta + \wt{\delta})}{a_3\sigma_K p_{\min}} \Bigl(1 + A_3\frac{\wt{\delta}}{r}\Bigr) b_r(u)} \label{pf:prop_isometry_5}
+& \overset{(v)}{\leq} \wt{r} \sqrt{\frac{1 + A_1(\theta + \wt{\delta})}{a_3\sigma_K p_{\min}} \Bigl(1 + A_3\frac{\wt{\delta}}{r}\Bigr) b_r(u)},\label{pf:prop_isometry_5}
 \end{align}
@@
 & \leq \frac{p_{\max}}{a_3p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)^2 \|\wt{\mc{P}}^{\star}u\|_{\wt{P}_n}^2 \\
-& = \frac{p_{\max}}{a_3p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)^2 \|u\|_{n}^2
-\end{align*}
-and plugging this estimate along with \eqref{pf:prop_isometry_5} back into \eqref{pf:prop_isometry_4}, upon choosing
-\begin{equation*}
-A_6 := \biggl(3\sqrt{\frac{2p_{\max}}{p_{\min}}} + 1\biggr)\sqrt{\frac{4}{a_3\sigma_Kp_{\min}}},~~A_7:=4A_1\frac{p_{\max}}{a_3p_{\min}}
-\end{equation*}
-we obtain part (2) of Proposition~\ref{prop:isometry}.
+& = \frac{p_{\max}}{a_3p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)^2 \|u\|_{n}^2.
+\end{align*}
+Plugging this estimate along with \eqref{pf:prop_isometry_5} back into \eqref{pf:prop_isometry_4}, we obtain part (2) of Proposition~\ref{prop:isometry}, upon choosing
+\begin{equation*}
+A_6 := \biggl(3\sqrt{\frac{2p_{\max}}{p_{\min}}} + 1\biggr)\sqrt{\frac{4}{a_3\sigma_Kp_{\min}}},~~A_7:=4A_1\frac{p_{\max}}{a_3p_{\min}}.
+\end{equation*}
 
@@
 	\label{eqn:empirical_norm_sobolev_1}
-	\norm{f}_{\Leb^2(\Xset)} \geq \frac{C_6 M}{\delta} \cdot \max\Bigl\{n^{-1/2},n^{-1/d}\Bigr\}
+	\norm{f}_{\Leb^2(\Xset)} \geq \frac{C_6 M}{\delta} \cdot \max\Bigl\{n^{-1/2},n^{-1/d}\Bigr\}.
 	\end{equation}
@@
 The nature of the inequalities we use depend on the value of $d$.
-In particular, we will use the following relationships between norms:
+In particular, we will use the following relationships between norms: for any $f \in H^1(\mc{X};M)$,
 \begin{equation*}
@@
 	\begin{equation*}
-	f^{\top} \Lap_{n,r} f \leq \frac{C}{\delta} n^2 r^{m + 2} |f|_{H^1(\Xset)}^2
+	f^{\top} \Lap_{n,r} f \leq \frac{C}{\delta} n^2 r^{m + 2} |f|_{H^1(\Xset)}^2.
 	\end{equation*}
@@
 	\begin{equation*}
-	\sqrt{\lambda_k(\Delta_P)}r + C(\theta + \wt{\delta}) \leq \frac{1}{2}
+	\sqrt{\lambda_k(\Delta_P)}r + C(\theta + \wt{\delta}) \leq \frac{1}{2},
 	\end{equation*} 
@@
 	\begin{equation*}
-	nr^{m+2} \lambda_k(\Delta_P) \biggl(1 - C\Bigl(r(\sqrt{\lambda_k(\Delta_P)} + 1) + \frac{\wt{\delta}}{r} + \theta\Bigr)\biggr)\leq \lambda_k(G_{n,r}) \leq nr^{m+2} \lambda_k(\Delta_P) \biggl(1 + C\Bigl(r(\sqrt{\lambda_k(\Delta_P)} + 1) + \frac{\wt{\delta}}{r} + \theta\Bigr)\biggr)
+	nr^{m+2} \lambda_k(\Delta_P) \biggl(1 - C\Bigl(r(\sqrt{\lambda_k(\Delta_P)} + 1) + \frac{\wt{\delta}}{r} + \theta\Bigr)\biggr)\leq \lambda_k(G_{n,r}) \leq nr^{m+2} \lambda_k(\Delta_P) \biggl(1 + C\Bigl(r(\sqrt{\lambda_k(\Delta_P)} + 1) + \frac{\wt{\delta}}{r} + \theta\Bigr)\biggr).
 	\end{equation*}
@@
 \label{eqn:neighborhood_eigenvalue_manifold}
-cB_{n,r}(k) \leq \lambda_k \leq CB_{n,r}(k),~~\textrm{for all $2 \leq k \leq n$}
+cB_{n,r}(k) \leq \lambda_k \leq CB_{n,r}(k),~~\textrm{for all $2 \leq k \leq n$}.
 \end{equation}
@@
 & = \biggl(\frac{C_2}{\delta} + \frac{40}{c_3^{2}}\biggr)M^{2d/(2+d)} n^{-2/(2 + d)} + \frac{10}{n} + \frac{10}{c_3^2}r^4M^{8/(2 + d)}n^{4/(2 + d)} \\
-& \leq \biggl(\frac{C_2}{\delta} + \frac{50}{c_3^{2}}\biggr)M^{2d/(2+d)} n^{-2/(2 + d)}
+& \leq \biggl(\frac{C_2}{\delta} + \frac{50}{c_3^{2}}\biggr)M^{2d/(2+d)} n^{-2/(2 + d)},
 \end{align*}
@@
 \paragraph{Proof of Theorem~\ref{thm:laplacian_smoothing_estimation2}.}
-We first establish that $\wh{f}$ achieves nearly-optimal rates when $d = 4$, and then prove the sub-optimal rates when $d > 4$.
+We first establish that $\wh{f}$ achieves nearly-optimal rates when $d = 4$, and then establish the claimed sub-optimal rates when $d > 4$.
 
@@
 \begin{equation*}
-\|\wh{f} - f_0\|_n^2 \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} +  \frac{10}{nc_3^2\wt{\rho}^2} +  \frac{10 \log n}{nc_3^2\wt{\rho}^2} + \frac{10r^4}{c_3^2\wt{\rho}^2}
+\|\wh{f} - f_0\|_n^2 \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} +  \frac{10}{nc_3^2\wt{\rho}^2} +  \frac{10 \log n}{nc_3^2\wt{\rho}^2} + \frac{10r^4}{c_3^2\wt{\rho}^2}.
 \end{equation*}
@@
 
-Once again continuing on from \eqref{pf:laplacian_smoothing_estimation1_1}, from Lemma~\ref{lem:variance_term_estimation} we have that,
-\begin{equation*}
-\|\wh{f} - f_0\|_n^2 \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{nc_3^{2}\wt{\rho}^{d/2}} +  \frac{10}{n^{4/d}\wt{\rho}^2c_3^2} + \frac{10r^4}{\wt{\rho}^2c_3^2}
+Once again continuing on from \eqref{pf:laplacian_smoothing_estimation1_1}, from Lemma~\ref{lem:variance_term_estimation} we have that
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 \leq \frac{C_2}{\delta} \wt{\rho} M^2 + \frac{10}{n} + \frac{10}{nc_3^{2}\wt{\rho}^{d/2}} +  \frac{10}{n^{4/d}\wt{\rho}^2c_3^2} + \frac{10r^4}{\wt{\rho}^2c_3^2}.
 \end{equation*}
@@
 \begin{equation*}
-\bigl(\wh{f}(x) - f_0(x)\bigr)^2 = \bigl(\wh{f}(X_i) - f_0(x)\bigr)^2 \leq 2\bigl(\wh{f}(X_i) - f_0(X_i)\bigr)^2 + 2\bigl(f_0(X_i) - f_0(x)\bigr)^2
+\bigl(\wh{f}(x) - f_0(x)\bigr)^2 = \bigl(\wh{f}(X_i) - f_0(x)\bigr)^2 \leq 2\bigl(\wh{f}(X_i) - f_0(X_i)\bigr)^2 + 2\bigl(f_0(X_i) - f_0(x)\bigr)^2.
 \end{equation*}
@@
 \int_{\Xset} \bigl(\wh{f}(x) - f_0(x)\bigr)^2 \,dx & \leq 2  \sum_{i = 1}^{n} \int_{V_i} \Bigl(\wh{f}(X_i) - f_0(X_i)\Bigr)^2 \,dx + 2 \sum_{i = 1}^{n} \int_{V_i} \Bigl(f_0(X_i) - f_0(x)\Bigr)^2 \dx \\
-& = 2 \sum_{i = 1}^{n} \vol(V_i) \Bigl(\wh{f}(X_i) - f_0(X_i)\Bigr)^2 + 2 \sum_{i = 1}^{n} \int_{V_i} \Bigl(f_0(X_i) - f_0(x)\Bigr)^2 \dx
+& = 2 \sum_{i = 1}^{n} \vol(V_i) \Bigl(\wh{f}(X_i) - f_0(X_i)\Bigr)^2 + 2 \sum_{i = 1}^{n} \int_{V_i} \Bigl(f_0(X_i) - f_0(x)\Bigr)^2 \dx,
 \end{align*}
@@
 \label{pf:out_of_sample_error_0}
-\|\wh{f} - f\|_{\Leb^2(\Xset)}^2 \leq 2 \sum_{i = 1}^{n} \vol(V_i) \Bigl(\wh{f}(X_i) - f_0(X_i)\Bigr)^2 + 2 M^2 \sum_{i = 1}^{n} \Bigl(\mathrm{diam}(V_i)\Bigr)^2
-\end{equation}
-where we write $\mathrm{diam}(V)$ for the diameter of a set $V$.
+\|\wh{f} - f\|_{\Leb^2(\Xset)}^2 \leq 2 \sum_{i = 1}^{n} \vol(V_i) \Bigl(\wh{f}(X_i) - f_0(X_i)\Bigr)^2 + 2 M^2 \sum_{i = 1}^{n} \Bigl(\mathrm{diam}(V_i)\Bigr)^2.
+\end{equation}
+Here we have written $\mathrm{diam}(V)$ for the diameter of a set $V$.
 
@@
 \begin{equation*}
-\Ebb\bigl[\wh{T}\bigr] = \mathbb{E}\bigl[f_0^2(X)\bigr] + 1 \geq p_{\min} \norm{f_0}_{\Leb^2(\Xset)}^2 + 1
+\Ebb\bigl[\wh{T}\bigr] = \mathbb{E}\bigl[f_0^2(X)\bigr] + 1 \geq p_{\min} \norm{f_0}_{\Leb^2(\Xset)}^2 + 1.
 \end{equation*}
@@
 	\label{eqn:paley_zygmund_pf1}
-	\mathbb{P}(Z > \lambda \mathbb{E}(Z^p)) \geq \left[(1 - \lambda^p) \frac{\mathbb{E}(Z^p)}{(\mathbb{E}(Z^q))^{p/q}}\right]^{\frac{q}{q - p}}
+	\mathbb{P}(Z > \lambda \mathbb{E}(Z^p)) \geq \left[(1 - \lambda^p) \frac{\mathbb{E}(Z^p)}{(\mathbb{E}(Z^q))^{p/q}}\right]^{\frac{q}{q - p}}.
 	\end{equation}
@@
 	\begin{equation*}
-	\Pbb\biggl(\Bigl|S_n - \mu\Bigr| \geq \delta \mu\biggr) \leq 2\exp\biggl(-\frac{\delta^2\mu}{3B^2}\biggr)
+	\Pbb\biggl(\Bigl|S_n - \mu\Bigr| \geq \delta \mu\biggr) \leq 2\exp\biggl(-\frac{\delta^2\mu}{3B^2}\biggr).
 	\end{equation*}
@@
 \label{pf:max_degree_1}
-\Pbb\biggl(\Bigl|D_{n,r}(x) - \Ebb\bigl[D_{n,r}(x)\bigr]\Bigr| \geq \delta \Ebb\bigl[D_{n,r}(x)\bigr]\biggr) \leq 2\exp\biggl(-\frac{\delta^2\Ebb\bigl[D_{n,r}(x)\bigr]}{3[K(0)]^2}\biggr)
-\end{equation}
-
+\Pbb\biggl(\Bigl|D_{n,r}(x) - \Ebb\bigl[D_{n,r}(x)\bigr]\Bigr| \geq \delta \Ebb\bigl[D_{n,r}(x)\bigr]\biggr) \leq 2\exp\biggl(-\frac{\delta^2\Ebb\bigl[D_{n,r}(x)\bigr]}{3[K(0)]^2}\biggr).
+\end{equation}
 Now we lower bound $\Ebb[D_{n,r}(x)]$ using the boundedness of the density $p$, and the fact that $\Xset$ has Lipschitz boundary:
@@
 & \geq n p_{\min} a_3 \int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx \\
-& \geq nr^d p_{\min}.
+& \geq nr^d p_{\min},
 \end{align*}
@@
 \begin{equation*}
-\Ebb\bigl[D_{n,r}(x)\bigr] \leq nr^{d} p_{\max}
+\Ebb\bigl[D_{n,r}(x)\bigr] \leq nr^{d} p_{\max},
 \end{equation*} 
@@
 \begin{equation*}
-\Pbb\biggl(D_{n,r}(x) \geq (1 + \delta) nr^d p_{\max}\biggr) \leq 2\exp\biggl(-\frac{\delta^2nr^da_0p_{\min}}{3[K(0)]^2}\biggr),
-\end{equation*}
-and applying a union bound, we get that
-\begin{equation*}
-\Pbb\biggl(\max_{i = 1,\ldots,n}D_{n,r}(X_i) \geq (1 + \delta) nr^d p_{\max}\biggr) \leq 2n\exp\biggl(-\frac{\delta^2nr^da_0p_{\min}}{3[K(0)]^2}\biggr);
-\end{equation*}
-taking $\delta = 1$ gives the claimed upper bound.
+\Pbb\biggl(D_{n,r}(x) \geq (1 + \delta) nr^d p_{\max}\biggr) \leq 2\exp\biggl(-\frac{\delta^2nr^da_0p_{\min}}{3[K(0)]^2}\biggr).
+\end{equation*}
+Applying a union bound, we get that
+\begin{equation*}
+\Pbb\biggl(\max_{i = 1,\ldots,n}D_{n,r}(X_i) \geq (1 + \delta) nr^d p_{\max}\biggr) \leq 2n\exp\biggl(-\frac{\delta^2nr^da_0p_{\min}}{3[K(0)]^2}\biggr),
+\end{equation*}
+and taking $\delta = 1$ gives the claimed upper bound.

### 2021-03-07 manuscripts/laplacian_smoothing/arXiv/appendix.tex
Responded to all but one of Siva's comments.

@@
 \end{align*}
-\sbcomment{General comment. Please punctuate math. I've done it till this point but please check everywhere.}
+\sbcomment{General comment. Please punctuate math. I've done it till this point but please check everywhere.} \\
+\agcomment{Resolved, I hope.} \\
 Writing $\norm{S\varepsilon}_2^2 = \sum_{k = 1}^{n} \lambda_k(S)^2 Z_k^2$, the claim follows from the result of \citet{laurent00} on concentration of $\chi^2$-random variables, which for completeness we restate in Lemma~\ref{lem:chi_square_bound}.
@@
 where the final inequality is satisfied because $S$ is a contraction.
-Plugging~\eqref{pf:linear_smoother_fixed_graph_testing2.5} and~\eqref{pf:linear_smoother_fixed_graph_testing3} back into~\eqref{pf:linear_smoother_fixed_graph_testing2} then gives the desired result.
-
-\sbcomment{I think the first equality should be the negative of what you wrote?}
-
+Plugging~\eqref{pf:linear_smoother_fixed_graph_testing2.5} and~\eqref{pf:linear_smoother_fixed_graph_testing3} back into~\eqref{pf:linear_smoother_fixed_graph_testing2} then gives the desired result.\\
+\sbcomment{I think the first equality should be the negative of what you wrote?} \\
 \agcomment{You were correct.
@@
 \end{theorem}
-We note that Theorem~\ref{thm:neighborhood_eigenvalue} is not a strong enough result to imply that $\lambda_{k}(G_{n,r})$ is a consistent estimate of $\lambda_k(\Delta_P)$, i.e.
+Before moving forward to the proofs of Lemma~\ref{lem:neighborhood_eigenvalue} and Theorem~\ref{thm:neighborhood_eigenvalue}, it is worth being clear about the differences between Theorem~\ref{thm:neighborhood_eigenvalue} and the results of \citet{burago2014,trillos2019,calder2019}.
+First of all, the reason we cannot directly use the results of these works in the proof of Lemma~\ref{lem:neighborhood_eigenvalue} is that they all assume the domain $\mc{X}$ is without boundary, whereas for our results in Section~\ref{sec:minimax_optimal_laplacian_smoothing} we instead assume $\mc{X}$ has a (Lipschitz smooth) boundary.
+Fortunately, in this setting the high-level strategy shared by \citet{burago2014,trillos2019,calder2019} can still be used---indeed we follow it closely, as we summarize in Section~\ref{subsec:neighbrhood_eigenvalue_pf}.
+However, many calculations need to be redone, in order to account for points $x$ which are on or sufficiently near to the boundary of $\mc{X}$.
+For completeness and ease of reading, we provide a self-contained proof of Theorem~\ref{thm:neighborhood_eigenvalue}, but we comment where appropriate on connections between the technical results we use in this proof, and those derived in \citet{burago2014,trillos2019,calder2019}.
+
+On the other hand, we should also point out that unlike the results of~\citet{burago2014,trillos2019,calder2019}, Theorem~\ref{thm:neighborhood_eigenvalue} does not imply that $\lambda_{k}(G_{n,r})$ is a consistent estimate of $\lambda_k(\Delta_P)$, i.e.
 it does not imply that $|(nr^{d + 2})^{-1}\lambda_{k}(G_{n,r}) - \lambda_k(\Delta_P)| \to 0$ as $n \to \infty, r \to 0$.
-This is contrast to the results of~\citet{burago2014,trillos2019,calder2019}, whose results do imply consistency when the domain is assumed to be a manifold without boundary.
-While a more subtle analysis might imply consistency in our setting, the conclusion of Theorem~\ref{thm:neighborhood_eigenvalue}---that $\lambda_k(G_{n,r})/(nr^{d + 2}\lambda_k(\Delta_P))$ is bounded above and below by constants---suffices for our purposes.
-
-The bulk of this section is devoted to the proof of Theorem~\ref{thm:neighborhood_eigenvalue}.
-First, however, we show that our regularity under our conditions on $p$ and $\Xset$, Lemma~\ref{lem:neighborhood_eigenvalue} is a simple consequence of Theorem~\ref{thm:neighborhood_eigenvalue}.
+The key difficulty in proving consistency when $\mc{X}$ has a boundary can be summarized as follows: while at points $x \in \mc{X}$ satisfying $B(x,r) \subseteq \mc{X}$, the graph Laplacian $L$ is a reasonable approximation of the operator $\Delta_P$, at points $x$ near the boundary $L$ is known to approximate a different operator altogether \citep{belkin2012}.
+This is reminiscent of the boundary effects present in the analysis of kernel smoothing.
+We believe a more subtle analysis might imply convergence of eigenvalues in this setting.
+However, the conclusion of Theorem~\ref{thm:neighborhood_eigenvalue}---that $\lambda_k(G_{n,r})/(nr^{d + 2}\lambda_k(\Delta_P))$ is bounded above and below by constants that do not depend on $k$---suffices for our purposes.
+
+\agcomment{Siva, I erred on the side of giving too much detail here, figuring it was easier for you/me to pare back if needed.}
+
+The bulk of the remainder of this section is devoted to the proof of Theorem~\ref{thm:neighborhood_eigenvalue}.
+First, however, we show that under our regularity conditions on $p$ and $\Xset$, Lemma~\ref{lem:neighborhood_eigenvalue} is a simple consequence of Theorem~\ref{thm:neighborhood_eigenvalue}.
 The link between the two is Weyl's Law.
@@
 \subsection{Proof of Theorem~\ref{thm:neighborhood_eigenvalue}}
+\label{subsec:neighbrhood_eigenvalue_pf}
 

### 2021-03-07 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
Changed proof of Lemma 4 in response to Siva's comment.
Responded to all but one of Siva's comments.

@@
 \newcommand{\sbcomment}[1]{{\bf{{\red{{SB --- #1}}}}}}
+\newcommand{\agcomment}[1]{{\bf{{\green{{AG --- #1}}}}}}
 
@@
 For sequences $a_n,b_n$, we write $a_n \lesssim b_n$ to mean $a_n \leq Cb_n$ for a constant $C>0$ and large enough $n$, and $a_n \asymp b_n$ to mean $a_n \lesssim b_n$ and $b_n \lesssim a_n$.
-Lastly, we use $a \wedge b = \max\{a,b\}$.
-\sbcomment{This is wrong, should check where it appears.}
+Lastly, we use $a \wedge b = \min\{a,b\}$.
+\\
+\sbcomment{This is wrong, should check where it appears.} \\
+\agcomment{Resolved.}
 
@@
 \label{eqn:graph_Laplacian}
-\Lap_{n,r} = D - W.
+\Lap = D - W.
 \end{equation}
-We add subscripts $n,r$ to $\Lap$ and $G$ to emphasize that they depend on the design points $X_1,\ldots,X_n$ and the radius $r$.
-To avoid any confusion: henceforth when we speak of \smash{$\wh{f}$} and \smash{$\wh{T}$}, we will always assume that they are fit using the neighborhood graph Laplacian \smash{$L = L_{n,r}$}.
-Writing its eigendecomposition as \smash{$L = \sum_{k = 1}^{n} \lambda_k v_k v_k^\top$}, we always assume, by convention, ordered eigenvalues $0 = \lambda_1 \leq \cdots \leq \lambda_n$, and unit-norm eigenvectors.
-%we will make reference to these objects later on.
-\sbcomment{Don't understand why you're switching between $L$ and $L_{n,r}$?}
+We write the eigendecomposition of $L$ as \smash{$L = \sum_{k = 1}^{n} \lambda_k v_k v_k^\top$}, and we always assume, by convention, ordered eigenvalues $0 = \lambda_1 \leq \cdots \leq \lambda_n$, and unit-norm eigenvectors.
+\\
+\sbcomment{Don't understand why you're switching between $L$ and $L_{n,r}$?} \\
+\agcomment{Resolved.}
 
@@
 It's fine, I wouldn't change anything because it would involve explaining a lot more, I'm just pointing this out.
-(Throughout we assume $M \geq n^{-1/2}$, as otherwise the minimax rate is just the parametric rate $n^{-1}$.) \sbcomment{I don't think the second part of this comment is correct the way you've defined the Sobolev class.}
+(Throughout we assume $M \geq n^{-1/2}$, as otherwise the trivial estimator $\wh{f} = 0$ achieves smaller error than the parametric rate $n^{-1}$, and the problem doesn't fit well within the nonparametric setup.)
+\\
+\sbcomment{I don't think the second part of this comment is correct the way you've defined the Sobolev class.} \\
+\agcomment{Resolved, as I suggested in my Slack post.}
 
@@
 In this paper, as in \citet{ingster2009}, we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\star} = 0$ and $\mc{F} = H^1(\Xset,M)$ is a first-order Sobolev ball.
-\sbcomment{Why isn't this without loss of generality?}
+\\
+\sbcomment{Why isn't this without loss of generality?}\\
+\agcomment{Don't know how to resolve.}
 
@@
 Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume the neighborhood graph \smash{$G_{n,r}$} is computed with a kernel $K$ satisfying \ref{asmp:kernel}.
-There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_estimation}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$},
+There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_estimation}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing} with \smash{$\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$},
 	with probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c (M^2n)^{d/(2 + d)})$, satisfies
@@
 	% AJG: Added.
-	\item The assumption $M \leq n^{1/d}$ ensures that the upper bound provided in the theorem is meaningful (i.e., ensures it is of most a constant order).
+	\item The assumption $M \leq n^{1/d}$ ensures that the upper bound provided in the theorem is meaningful (i.e., ensures it is of at most a constant order).
 	\item The lower bound on $r$ imposed in condition \ref{asmp:ls_kernel_radius_estimation} is compatible with practice, where by far the most common choice of radius is the connectivity threshold \smash{$r \asymp (\log(n)/n)^{1/d}$}, which makes \smash{$G_{n,r}$} as sparse as possible while still being connected, for maximum computational efficiency.
@@
 We leave analysis of the Sobolev case to future work.
-	\item When $f_0$ is Lipschitz smooth, we can also replace the factor of $\delta$ in the high probability bound by the sharper $\delta^2/n$. \sbcomment{Too mysterious a comment. Why is this sharper?}
+	\item When $f_0$ is Lipschitz smooth, we can also replace the factor of $\delta$ in the high probability bound by a factor of $\delta^2/n$, which is always smaller than $\delta$ when $\delta \in (0,1)$.\\
+	\sbcomment{Too mysterious a comment. Why is this sharper?} \\
+	\agcomment{Attempted to resolve.} 
 \end{itemize}
@@
 \end{theorem}
-\sbcomment{I removed some negative sign above. Pls check.}
+\sbcomment{I removed some negative sign above. Pls check.}\\
+\agcomment{Good, thanks.}\\
 This mirrors the conclusions of \citet{sadhanala16} who investigate estimation rates of Laplacian smoothing over the $d$-dimensional grid graph.
@@
 \end{enumerate}
-We will also require that the radius of the graph Sobolev class \sbcomment{What does the class have to do with graphs?} not be too large.
+We will also require that the radius of the Sobolev class not be too large. \\
+\sbcomment{What does the class have to do with graphs?}  \\
+\agcomment{Resolved.} \\
 Precisely, we will require $M \leq M_{\max}(d)$, where we define
@@
 Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume \smash{$G_{n,r}$} is computed with a kernel $K$ satisfying \ref{asmp:kernel}.
-There are constants $N,C,c>0$ such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_testing}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$}, satisfies the following: for all $b \geq 1$, and all $f_0$ such that
+There are constants $N,C,c>0$ such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_testing}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$}, satisfies the following: for all $b \geq 1$, and all $f_0$ such that
 	\begin{equation}
@@
 \end{equation*}
-and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies e.g., Dirichlet boundary conditions.} To be clear, this argument does not formally imply that $\wh{f}$ and $\wt{f}$ are close \sbcomment{Should remind people what these are, or refer to eqns where they are defined.} (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of continuous functions with weak derivatives in $\Leb^2(\Xset)$).
+and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies e.g., Dirichlet boundary conditions.} To be clear, this argument does not formally imply that the Laplacian eigenmaps estimator $\wh{f}$ and the thin-plate spline estimator $\wt{f}$ are close (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of continuous functions with weak derivatives in $\Leb^2(\Xset)$).
 But it does seem to suggest that the two estimators should behave somewhat similarly.
@@
 Lemma~\ref{lem:neighborhood_eigenvalue} leads to a sufficiently tight upper bound on the variance term whenever the radius $r$ is sufficiently small; precisely, when $r$ is upper bounded as in \ref{asmp:ls_kernel_radius_estimation} for estimation, or \ref{asmp:ls_kernel_radius_testing} for testing.
-The parameter $\rho$ is then tuned to minimize the sum of bias and variance, in the usual manner.
-\sbcomment{Should probably say something about the rest of the proof, i.e. is it straightforward to put these together to get the two Theorems?}
+The parameter $\rho$ is then chosen to minimize the sum of these upper bounds on bias and variance, as usual, and some straightforward calculations gives Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing}.\\
+\sbcomment{Should probably say something about the rest of the proof, i.e. is it straightforward to put these together to get the two Theorems?} \\
+\agcomment{Attempted to resolve.}
 
@@
 	As in Theorem \ref{thm:laplacian_smoothing_testing}, but where $\Xset \subseteq \Rd$ is a manifold with intrinsic dimension $m < 4$, $M \leq M_{\max}(m)$, and the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}.
-	There are constants $N,C,c>0$ such that for any $n \geq N$, and any $r$ as in \ref{asmp:ls_kernel_radius_testing_manifold}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$}, satisfies the following: for all $b \geq 1$, and $f_0$ such that
+	There are constants $N,C,c>0$ such that for any $n \geq N$, and any $r$ as in \ref{asmp:ls_kernel_radius_testing_manifold}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$}, satisfies the following: for all $b \geq 1$, and $f_0$ such that
 	\begin{equation}
@@
 In practice, it is more common to use a $k$-nearest-neighbor (kNN) graph than a neighborhood graph, due to the guaranteed connectivity and sparsity of the former; we suspect that by building on the work of \citet{calder2019}, one can show that our main results all hold under the kNN graph as well.
-In another direction, one can also generalize Laplacian smoothing by replacing the penalty \smash{$f^\top \Lap_{n,r} f$} with \smash{$f^\top \Lap_{n,r}^k f$}, for an integer $k > 1$.
-The hope is that this would then achieve minimax optimal rates over the higher-order Sobolev class $H^k(\Xset)$.
+In another direction, one can also generalize Laplacian smoothing by replacing the penalty \smash{$f^\top \Lap_{n,r} f$} with \smash{$f^\top \Lap_{n,r}^{s} f$}, for an integer $s > 1$.
+The hope is that this would then achieve minimax optimal rates over the higher-order Sobolev class $H^s(\Xset)$.
 % In the very special case of $k = 2$ and $d \leq 2$, we can show that this is indeed the case, but the general story--for all combinations of $k$ and $d$---remains beyond our reach.
-\sbcomment{Not a huge fan of using $k$ in adjacent lines to refer to completely different things.}
+\sbcomment{Not a huge fan of using $k$ in adjacent lines to refer to completely different things.}\\
+\agcomment{Resolved.}
 

### 2021-03-08 manuscripts/laplacian_smoothing/arXiv/appendix.tex
camera ready changes

@@
 \end{equation*}
----where in the above the maximum is over all indices $k$ such that $\lambda_k > 0$, and the last inequality follows from the basic algebraic identity $1 - 1/(1 + x)^2 \leq 2 x$ for any $x > 0$---we deduce that
+---where in the above the maximum is over all indices $k$ such that $\lambda_k > 0$, and the last inequality follows from the basic algebraic identity $1 - 1/(1 + x)^2 \leq 2 x$ for any $x > 0$ \sbcomment{You've used this before, so maybe move this comment earlier?}---we deduce that
 \begin{equation*}
@@
 \end{equation*}
-To avoid confusion, in this section we write $\lambda_k(G_{n,r})$ for the $k$th smallest eigenvalue of the graph Laplacian matrix $\Lap_{n,r}$ and $\lambda_k(\Delta_P)$ for the $k$th smallest eigenvalue of $\Delta_P$ \footnote{Under the assumptions~\ref{asmp:domain} and~\ref{asmp:density}, the operator $\Delta_P$ has a discrete spectrum; see \citet{garciatrillos18} for more details.}.
+To avoid confusion, in this section we write $\lambda_k(G_{n,r})$ for the $k$th smallest eigenvalue of the graph Laplacian matrix $\Lap_{n,r}$ and $\lambda_k(\Delta_P)$ for the $k$th smallest eigenvalue of $\Delta_P$ \footnote{Under the assumptions~\ref{asmp:domain} and~\ref{asmp:density}, the operator $\Delta_P$ has a discrete spectrum; see \citet{garciatrillos18} \sbcomment{Fix this reference, typos.} for more details.}.
 Some other notation: throughout this section, we will write $A, A_0, A_1,\ldots$ and $a,a_0,a_1,\ldots$ for constants which may depend on $\Xset$, $d$, $K$, and $p$, but do not depend on $n$; we keep track of all such constants explicitly in our proofs.
@@
 Fortunately, in this setting the high-level strategy shared by \citet{burago2014,trillos2019,calder2019} can still be used---indeed we follow it closely, as we summarize in Section~\ref{subsec:neighbrhood_eigenvalue_pf}.
-However, many calculations need to be redone, in order to account for points $x$ which are on or sufficiently near to the boundary of $\mc{X}$.
+However, many calculations need to be redone, in order to account for points $x$ which are on or sufficiently close to the boundary of $\mc{X}$.
 For completeness and ease of reading, we provide a self-contained proof of Theorem~\ref{thm:neighborhood_eigenvalue}, but we comment where appropriate on connections between the technical results we use in this proof, and those derived in \citet{burago2014,trillos2019,calder2019}.
@@
 \agcomment{Siva, I erred on the side of giving too much detail here, figuring it was easier for you/me to pare back if needed.}
+\sbcomment{It is extremely insightful actually. I would consider including some of the second paragraph in the main paper too.}
 
@@
 \end{equation*}
-Let us verify that $\lambda_{\ell_{\star}}(\Delta_P)$ satisfies the condition~\ref{eqn:neighborhood_eigenvalue_1} of Theorem~\ref{thm:neighborhood_eigenvalue}.
+Let us verify that $\lambda_{\ell_{\star}}(\Delta_P)$ satisfies the condition~\eqref{eqn:neighborhood_eigenvalue_1} of Theorem~\ref{thm:neighborhood_eigenvalue}.
 Setting $c_0 := 1/(2^{1/d}4A_2^{1/2})$, the assumed upper bound on the radius $r \leq c_0$ guarantees that $\ell_{\star} \geq 2$.
@@
 
-\begin{proposition}[\textbf{c.f. Proposition 4.1 of \citet{calder2019}}]
+\begin{proposition}[\textbf{cf. Proposition 4.1 of \citet{calder2019}}]
 	\label{prop:dirichlet_energies}
@@
 
-\begin{proposition}[\textbf{c.f. Proposition 4.2 of \citet{calder2019}}]
+\begin{proposition}[\textbf{cf. Proposition 4.2 of \citet{calder2019}}]
 	\label{prop:isometry}
@@
 In Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, we establish the reverse inequality.
-\begin{lemma}[\textbf{c.f. Lemma 9 of \citet{trillos2019}, Lemma 5.5 of \citet{burago2014}}]
+\begin{lemma}[\textbf{cf. Lemma 9 of \citet{trillos2019}, Lemma 5.5 of \citet{burago2014}}]
 	\label{lem:first_order_graph_sobolev_seminorm_expected_lb}
@@
 We will use Lemma~\ref{lem:poincare} in the proof of Proposition~\ref{prop:isometry}. 
-\begin{lemma}[\textbf{c.f. Lemma 3.4 of \citet{burago2014}}]
+\begin{lemma}[\textbf{cf. Lemma 3.4 of \citet{burago2014}}]
 	\label{lem:poincare}
@@
 Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized} relates the graph Sobolev semi-norm $b_r(\wt{\mc{P}}f)$ to the non-local energy $\wt{E}_r(f)$. 
-\begin{lemma}[\textbf{c.f. Lemma 13 of \citet{trillos2019}, Lemma 4.3 of \citet{burago2014}}]
+\begin{lemma}[\textbf{cf. Lemma 13 of \citet{trillos2019}, Lemma 4.3 of \citet{burago2014}}]
 	\label{lem:first_order_graph_sobolev_seminorm_discretized}
@@
 In Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}, we establish the reverse of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}. 
-\begin{lemma}[\textbf{c.f. Lemma 14 of \citet{trillos2019}}]
+\begin{lemma}[\textbf{cf. Lemma 14 of \citet{trillos2019}}]
 	\label{lem:first_order_graph_sobolev_seminorm_discretized_lb}

### 2021-03-08 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
camera ready changes

@@
 \\
-\sbcomment{This is wrong, should check where it appears.} \\
-\agcomment{Resolved.}
+%\sbcomment{This is wrong, should check where it appears.} \\
+%\agcomment{Resolved.}
 
@@
 \\
-\sbcomment{Don't understand why you're switching between $L$ and $L_{n,r}$?} \\
-\agcomment{Resolved.}
+%\sbcomment{Don't understand why you're switching between $L$ and $L_{n,r}$?} \\
+%\agcomment{Resolved.}
 
@@
 \\
-\sbcomment{I don't think the second part of this comment is correct the way you've defined the Sobolev class.} \\
-\agcomment{Resolved, as I suggested in my Slack post.}
+%\sbcomment{I don't think the second part of this comment is correct the way you've defined the Sobolev class.} \\
+%\agcomment{Resolved, as I suggested in my Slack post.}
 
@@
 In this paper, as in \citet{ingster2009}, we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\star} = 0$ and $\mc{F} = H^1(\Xset,M)$ is a first-order Sobolev ball.
-\\
-\sbcomment{Why isn't this without loss of generality?}\\
-\agcomment{Don't know how to resolve.}
+This is without loss of generality since our test statistic and its analysis are easily modified to handle the case when $f_0^{\star}$ is not $0$, by simply subtracting $f_0^{\star}(X_i)$ from each observation $Y_i$.
+
+%\sbcomment{Why isn't this without loss of generality?}\\
+%\agcomment{Don't know how to resolve.}
 
@@
 \end{theorem}
-\sbcomment{I removed some negative sign above. Pls check.}\\
-\agcomment{Good, thanks.}\\
+%\sbcomment{I removed some negative sign above. Pls check.}\\
+%\agcomment{Good, thanks.}\\
 This mirrors the conclusions of \citet{sadhanala16} who investigate estimation rates of Laplacian smoothing over the $d$-dimensional grid graph.
@@
 We will also require that the radius of the Sobolev class not be too large. \\
-\sbcomment{What does the class have to do with graphs?}  \\
-\agcomment{Resolved.} \\
+%\sbcomment{What does the class have to do with graphs?}  \\
+%\agcomment{Resolved.} \\
 Precisely, we will require $M \leq M_{\max}(d)$, where we define
@@
 But in practice we would not recommend using \smash{$\wh{t}_b$} anyway, and would instead we make the standard recommendation to calibrate via a permutation test \citep{hoeffding1952}.
+Recent work \cite{kim2020minimax}, has shown that in a variety of closely related settings, calibration of a test statistic via the permutation test often retains minimax-optimal power, and we expect similar results to hold for the Laplacian smoothing-based test statistic.
+
 \end{itemize}
@@
 Lemma~\ref{lem:neighborhood_eigenvalue} leads to a sufficiently tight upper bound on the variance term whenever the radius $r$ is sufficiently small; precisely, when $r$ is upper bounded as in \ref{asmp:ls_kernel_radius_estimation} for estimation, or \ref{asmp:ls_kernel_radius_testing} for testing.
-The parameter $\rho$ is then chosen to minimize the sum of these upper bounds on bias and variance, as usual, and some straightforward calculations gives Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing}.\\
+The parameter $\rho$ is then chosen to minimize the sum of these upper bounds on bias and variance, as usual, and some straightforward calculations give Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing}.\\
 \sbcomment{Should probably say something about the rest of the proof, i.e. is it straightforward to put these together to get the two Theorems?} \\
@@
 Under such an assumption, it is known \citep{bickel2007,ariascastro2018} that the optimal rates over $H^1(\Xset)$ are now $n^{-2/(2 + m)}$ (for estimation) and $n^{-4/(4 + m)}$ (for testing), which are much faster than the full-dimensional error rates when $m \ll d$.
+\sbcomment{I might point out that the Arias-Castro result is for density testing over Holder spaces, and that technically rates for regression testing on manifolds for Sobolev spaces are your contribution.}
 
@@
 
-On the other hand, a theory has been developed \citep{belkin03,belkin05,niyogi2013} establishing that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+On the other hand, a theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2011minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
 We contribute to this line of work by showing that under the manifold hypothesis, Laplacian smoothing achieves the tighter minimax rates over $H^1(\Xset)$.
@@
 % In the very special case of $k = 2$ and $d \leq 2$, we can show that this is indeed the case, but the general story--for all combinations of $k$ and $d$---remains beyond our reach.
-\sbcomment{Not a huge fan of using $k$ in adjacent lines to refer to completely different things.}\\
-\agcomment{Resolved.}
+%\sbcomment{Not a huge fan of using $k$ in adjacent lines to refer to completely different things.}\\
+%\agcomment{Resolved.}
 

### 2021-03-08 manuscripts/laplacian_smoothing/arXiv/appendix.tex
camera ready attempt 2

@@
 \end{align*}
-\sbcomment{General comment. Please punctuate math. I've done it till this point but please check everywhere.} \\
-\agcomment{Resolved, I hope.} \\
+%\sbcomment{General comment. Please punctuate math. I've done it till this point but please check everywhere.} \\
+%\agcomment{Resolved, I hope.} \\
 Writing $\norm{S\varepsilon}_2^2 = \sum_{k = 1}^{n} \lambda_k(S)^2 Z_k^2$, the claim follows from the result of \citet{laurent00} on concentration of $\chi^2$-random variables, which for completeness we restate in Lemma~\ref{lem:chi_square_bound}.
@@
 Plugging~\eqref{pf:linear_smoother_fixed_graph_testing2.5} and~\eqref{pf:linear_smoother_fixed_graph_testing3} back into~\eqref{pf:linear_smoother_fixed_graph_testing2} then gives the desired result.\\
-\sbcomment{I think the first equality should be the negative of what you wrote?} \\
-\agcomment{You were correct.
+%\sbcomment{I think the first equality should be the negative of what you wrote?} \\
+%\agcomment{You were correct.
 However I also rewrote the proof a bit to make it (hopefully) easier to follow.

### 2021-03-08 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
camera ready attempt 2

@@
 	\item When $f_0$ is Lipschitz smooth, we can also replace the factor of $\delta$ in the high probability bound by a factor of $\delta^2/n$, which is always smaller than $\delta$ when $\delta \in (0,1)$.\\
-	\sbcomment{Too mysterious a comment. Why is this sharper?} \\
-	\agcomment{Attempted to resolve.} 
+	%\sbcomment{Too mysterious a comment. Why is this sharper?} \\
+	%\agcomment{Attempted to resolve.} 
 \end{itemize}
@@
 The parameter $\rho$ is then chosen to minimize the sum of these upper bounds on bias and variance, as usual, and some straightforward calculations give Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing}.\\
-\sbcomment{Should probably say something about the rest of the proof, i.e. is it straightforward to put these together to get the two Theorems?} \\
-\agcomment{Attempted to resolve.}
+%\sbcomment{Should probably say something about the rest of the proof, i.e. is it straightforward to put these together to get the two Theorems?} \\
+%\agcomment{Attempted to resolve.}
 
@@
 
-On the other hand, a theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2011minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+On the other hand, a theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
 We contribute to this line of work by showing that under the manifold hypothesis, Laplacian smoothing achieves the tighter minimax rates over $H^1(\Xset)$.
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{unsrtnat}
 \bibliography{../../../graph_regression_bibliography}

### 2021-03-08 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Most of the way through section on out-of-sample extension for Laplacian Eigenmaps paper.

@@
 \textbf{(2) Zero-trace and periodic conditions.}
-When $s \geq 1$, our results regarding Laplacian eigenmaps will require some boundary assumptions on $f_0$.
+When $s \geq 1$, our results regarding Laplacian eigenmaps will require that $f_0$ satisfy some boundary assumptions.
 In particular, we will assume that $f_0$ belongs to the set of zero trace Sobolev functions $H_0^{s}(P)$, meaning \textcolor{red}{roughly} that $f_0 \in H^s(P)$ additionally satisfies,
@@
 \end{proposition}
-By Weyl's Law, $c k^{2/d} \leq \lambda_{k}(\Delta_P) \leq Ck^{2/d}$, and as a result $\lambda_{K} \geq C\lambda_{K}(\Delta_P)$ as long as $r \geq \kappa^{-1/d}$.
+By Weyl's Law, there exist constants $c_1$ and $C_1$ which do not depend on $k$, such that $c_1 k^{2/d} \leq \lambda_{k}(\Delta_P) \leq C_1 k^{2/d}$ for all $k \in \mathbb{N}$.
+As a result, $\lambda_{K} \geq c \lambda_{K}(\Delta_P)$ as long as $r \geq \kappa^{-1/d}$.
 
@@
 	\item \textbf{Restriction on number of derivatives.} Unlike in the full-dimensional case, our upper bound on the estimation error of $\wh{f}$ matches the minimax rate $n^{-2s/(2s + m)}$ only when $s \leq 4$.
+We now explain this discrepancy.
 At a high level, thinking of the graph $G$ as an estimate of the manifold $\mc{X}$, we incur some error by using Euclidean distance rather than geodesic distance to form the edges of $G$.
-By contrast, in the full-dimensional setting the Euclidean metric exactly coincides with the geodesic distance for all points $x,z \in \mc{X}$ that are sufficiently close to each other and far from the boundary of $\mc{X}$).
-This extra error incurred in the manifold setting dominates when $s > 4$.
-We give a more detailed explanation of this phenomenon in Appendix~\ref{sec:manifold_proofs}.
+In contrast, in the full-dimensional setting the Euclidean metric exactly coincides with the geodesic distance for all points $x,z \in \mc{X}$ that are sufficiently close to each other and far from the boundary of $\mc{X}$.
+This extra error incurred in the manifold setting by using the ``wrong distance'' dominates when $s > 4$.
+As this explanation suggests, by building $G$ using the geodesic distance one could avoid this error, and would obtain superior rates of convergence.
+However this is not an option for us, as we assume $\mc{X}$---and in particular its geodesics---are unknown.
 	\item \textbf{Comparison with traditional spectral series estimators.} Traditional spectral series estimators achieve the minimax rate for all values of $s$ and $m$.
-On the other hand, as we have already discussed, these estimators assume a priori knowledge of the domain $\mc{X}$, which is arguably a particularly unrealistic assumption when $\mc{X}$ is assumed to be a manifold.
+On the other hand, as we have already discussed, these estimators assume a priori knowledge of the domain $\mc{X}$, which is a particularly unrealistic assumption when $\mc{X}$ is assumed to be a manifold.
 It is not clear whether this gap between spectral series and Laplacian eigenmaps estimators---or more generally, between estimators which assume the manifold is known, and those which do not---is real, or a product of loose upper bounds.
@@
 \label{sec:out_of_sample}
-To summarize, Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity} show that $\wh{f}$ is a minimax optimal estimator over Sobolev spaces.
+Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity} show that $\wh{f}$ is a minimax optimal estimator over Sobolev spaces.
 However, as mentioned previously we have measured loss \emph{in-sample}---that is, measured in $\Leb^2(P_n)$ norm---whereas \emph{out-of-sample} error---error measured in $L^2(P)$ norm---is the more typical metric in the random design setup.
@@
 \paragraph{Extension by kernel smoothing.}
-We now formally define our approach to extending by kernel smoothing.
+We now formally define our approach to extension by kernel smoothing.
 For a kernel function $\psi(\cdot): [0,\infty) \to (-\infty,+\infty)$, bandwidth $h > 0$, and a distribution $Q$, the \emph{Nadaraya-Watson kernel smoother} $T_{\varepsilon,Q}$ is given by
@@
 
-The estimated function $\wt{f}$ we will consider is a kernel smoother passed over $\wh{f}$, meaning
+The estimated function $\wt{f} \in L^2(P)$ we will consider is a kernel smoother passed over $\wh{f}$, meaning
 \begin{equation}
@@
 It is useful to make a few comments about $\wt{f}$.
-First of all, while in principle~\eqref{eqn:kernel_smoother_laplacian_eigenmaps} makes sense for any $x \in \Rd$, we will always think of $\wt{f}$ as a function on $\mc{X}$.
+First of all, while in principle~\eqref{eqn:kernel_smoother_laplacian_eigenmaps} makes sense at any point $x \in \Rd$, we will always think of $\wt{f}$ as a function mapping $\mc{X}$ to $\Reals$.
 Additionally, note we may use the same kernel for smoothing as was used to construct the graph, i.e.
@@
 In Lemma~\ref{lem:kernel_smoothing_laplacian_eigenmaps}, we consider an arbitrary estimator $\wc{f} \in L^2(P_n)$.
-We show that the out-of-sample error $\|T_{n,h}\wc{f} - f_0\|_P^2$ can be upper bounded by three terms--- (a constant times) the in-sample error of $\wc{f} - f_0$, and bias and variance terms associated with a noiseless version of kernel smoothing.
+We show that the out-of-sample error $\|T_{n,h}\wc{f} - f_0\|_P^2$ can be upper bounded by three terms--- (a constant times) the in-sample error of $\wc{f} - f_0$, and variance and bias terms associated with a noiseless version of kernel smoothing.
 We shall assume the following conditions on $\psi$.
@@
 	\label{asmp:kernel}
-	The kernel function $\psi$ is supported on a subset of $[0,1]$, with $\psi(1/2) > 0$.
-Additionally, $\psi$ is Lipschitz continuous on $[0,1]$.
-Finally, for convenience, we will assume that $\psi$ is normalized such that
+	The kernel function $\psi$ is supported on a subset of $[0,1]$. Additionally, $\psi$ is Lipschitz continuous on $[0,1]$, and is normalized so that
 	\begin{equation*}
@@
 	\label{lem:kernel_smoothing_laplacian_eigenmaps}
-	Suppose $\wc{f} \in L^2(P_n)$, $f_0 \in H^1(\mc{X})$ and $p \in C^1(\mc{X})$. Under Assumptions~\ref{asmp:domain}, \ref{asmp:density}, and~\ref{asmp:kernel}, it holds that
+	Suppose $\wc{f} \in L^2(P_n)$, $f_0 \in H^1(P)$ and $p \in C^1(P)$. Under Assumptions~\ref{asmp:domain}, \ref{asmp:density}, and~\ref{asmp:kernel}, it holds that
 	\begin{equation}
@@
 Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data, by a factor of $h^2$.
-The bias term, on the other hand, is standard (at least when $f_0$ is assumed to be H\"{o}lder smooth), and can be tightly bounded using Taylor expansion when $\psi$ is an order-$s$ kernel.
-Recall that $\psi$ is an order-$s$ kernel if
+The bias term, on the other hand, is standard.
+A standard analysis shows that the $\|T_{h,P}f_0 - f_0\|_P^2 \lesssim \varepsilon^{2s}$ when $\psi$ is an order-$s$ kernel, meaning
 \begin{equation*}
@@
 		\end{equation*}
-		\item If $f_0 \in H_0^{s}(P)$, $p \in C^{r}(\mc{X})$ for $r = s - 1$, and $\psi$ is an order-$s$ kernel, it follows that
+		\item If $f_0 \in H_0^{s}(P)$, $p \in C^{s}(\mc{X})$, and $\psi$ is an order-$s$ kernel, it follows that
 		\begin{equation*}
-		\|T_{h,P}f_0 - f_0\|_P^2 \leq C h^{2s} |f|_{H^s(\mc{X})}^2.
+		\|T_{h,P}f_0 - f_0\|_P^2 \leq C h^{2s} |f|_{H^s(P)}^2.
 		\end{equation*}
@@
 \end{lemma}
-Choosing $h \asymp n^{-1/(2r + d)}$ for $r = s - 1$ balances the kernel smoothing bias and variance terms in~\eqref{eqn:kernel_smoothing_laplacian_eigenmaps}.
-Then, applying the bounds in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} (when $s = 1$) or Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} (when $s > 1$), we conclude that $\wt{f}$ is an optimal estimator, error measured in $L^2(P)$ norm.
+Choosing $h \asymp n^{-1/(2(s - 1) + d)}$ balances the kernel smoothing bias and variance terms in~\eqref{eqn:kernel_smoothing_laplacian_eigenmaps}.
+Then, applying the bounds in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} (when $s = 1$) or Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} (when $s > 1$), we conclude that $T_{h,n}\wh{f}$ is an optimal estimator, with respect to error measured in $L^2(P)$ norm.
 
+\begin{theorem}
+	\label{thm:laplacian_eigenmaps_estimation_out_of_sample}
+	Suppose we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to~\eqref{eqn:model}.
+Suppose that for some $s \in \mathbb{N}$, the regression function $f_0 \in H^s(P;M)$ (if $s = 1$) or $H_0^s(P;M)$ (if $s > 1$).
+Suppose also that the distribution $P$ has support $\mc{X}$ which satisfies~\ref{asmp:domain} and density $p \in C^s(\mc{X};M)$ which satisfies~\ref{asmp:density}.
+Then if the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_ho}, then the out-of-sample extension $T_{h,n}\wh{f}$ computed with bandwidth $h \asymp n^{-1/(2(s - 1) + d)}$ satisfies
+	\begin{equation*}
+	\|T_{h,n}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^{2d/(2s + d)}n^{-2s/(2s + d)},
+	\end{equation*}
+	with probability at least $1 - \delta - Cn\exp\{-Cn\varepsilon^d\} - Ch^d\exp\{-Cnh^d\}$. 
+\end{theorem}
 Some remarks:
@@
 	\item \textbf{Tuning parameters by sample-splitting.} 
-	\item \textbf{Other extensions.} Comment that there does not exist any canonical way to extend Laplacian based estimators such as $\wh{\theta}$ to an ambient domain.
-Point out that there exist various proposals to extend eigenvectors in the literature, such as $1$-nearest-neighbor regression (piecewise constant extrapolation over Voronoi cells), convolution with a specialized kernel, or Nystr\"{o}m extension.
-Explain that each of these is designed for a highly specialized purpose, such as suitability for mathematical analysis or computational efficiency, and in any case are suggested only for eigenvectors rather than a regression estimate.
-State explicitly that we choose kernel smoothing for its simplicity, ubiquity, and strong theoretical properties under our assumptions, but that considerations of extensions for specifically tailored to Laplacian based estimators could be of interest.
+	\item \textbf{Other extensions.} We consider kernel smoothing because it is a simple and intuitive procedure that does not rely on knowledge of the domain $\mc{X}$ or distribution $P$.
+There exist (many) other approaches to extending (or interpolating) a function $f$ based on the evaluations $\{f(X_1),\ldots,f(X_n)\}$, and we now comment on a few of these.
+One approach is to perform minimum norm interpolation, where one defines a normed space $(\mc{F},\|\cdot\|_{\mc{F}})$ and then solves the optimization problem
+	\begin{equation*}
+	\min_{u \in \mc{F}} \|u\|_{\mc{F}},~~\textrm{such that}~~ u(X_i) =  f(X_i)~~\textrm{for $i = 1,\ldots,n$.}
+	\end{equation*}
+	A particularly popular version of this general approach is when $\mc{F}$ is an RKHS, which encompasses thin-plate spline interpolation $\mc{F} = H^s(\mc{X})$ for $s > d/2$ as a special case.
+This approach works well when $f$ is close to a function $u \in \mc{F}$ with reasonably small norm, which holds true by assumption  but as already discussed when $2s \leq d$ the Sobolev space $H^s(P)$ is not a Reproducing Kernel Hilbert Space.
+Another
 	\item \textbf{Comparison with kernel smoothing.} 
@@
 
-\subsection{Why not just do kernel smoothing?}
-On the other hand, the allusion to kernel smoothing raises a very natural question: why should we prefer to first compute $\wh{\theta}$ and then smooth it out, as opposed to simply performing kernel smoothing over the original responses $Y$?
+\paragraph{Why not just kernel smoothing?}
+On the other hand, the allusion to kernel smoothing raises a very natural question: why should we prefer to first compute $\wh{f}$ and then smooth it out, as opposed to simply performing kernel smoothing over the original responses $Y$, that?
 

### 2021-03-08 notes/weekly_notes/12.24.20.NotesForWeek.tex
Most of the way through section on out-of-sample extension for Laplacian Eigenmaps paper.

@@
 
+Recall the notation $d_P(x) = \int \eta(\|z - x\|/\varepsilon) \,dP(z)$, $d_n(x) = d_{P_n}(x)$. Recall also that $K(\cdot) = |\eta(\cdot)|^2$. 
 \begin{lemma}
 	\label{lem:measure_of_ball_estimate}
-	With probability at least $1 - C \varepsilon^{-d}\exp\{-Cn\varepsilon^d\}$, it holds that
+	Let $\eta$ be a kernel compactly supported and Lipschitz on $[0,1]$. There exists a constant $C$ that does not depend on $\varepsilon$ or $\eta$, such that
 	\begin{equation*}
@@
 	\begin{equation*}
-	\sup_{x \in \Omega}|d_n(x;K^{1/2}) - d_P(x;K^{1/2})| \leq C \sqrt{\frac{\varepsilon^d}{n}}
+	\sup_{x \in \Omega}|d_n(x;K^{1/2}) - d_P(x;K^{1/2})| \leq C \sqrt{\frac{\varepsilon^d}{n}},
 	\end{equation*}
+	with probability at least $1 - \textcolor{red}{(?)}\exp\{-Cn\varepsilon^d\}$. 
 \end{lemma}
@@
 \begin{proof}
-	\textcolor{red}{(TODO)}
+	Since $\eta$ is compactly supported on $[0,1]$, a standard computation implies that since for any $x \in \mc{X}$,
+	\begin{equation*}
+	\Var\bigl[d_n(x)\bigl] = \frac{1}{n} \Var[\eta(\|X - x\|/\varepsilon)] \leq \frac{1}{n}\|\eta\|_{\infty}^2 \nu_d \varepsilon^d,
+	\end{equation*}
+	where we note that $\|\eta\|_{\infty} < \infty$ follows from the assumption that $\eta$ is Lipschitz. It follows from Bernstein's inequality that 
+	\begin{equation*}
+	\Pbb\biggl(|d_n(x) - d_P(x)| \leq C \sqrt{\frac{\varepsilon^d}{n}} \biggr) \leq 2\exp\biggl\{-\frac{t^2 \varepsilon^d}{C(\varepsilon^d + \delta \sqrt{\varepsilon^d/n})}\biggr\}.
+	\end{equation*}
+	for $C = \|\eta\|_{\infty} \nu_d$. \textcolor{red}{(...)}
 \end{proof}

### 2021-03-08 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
Separated symbols for type-I vs. type-II error.

@@
 (Throughout we assume $M \geq n^{-1/2}$, as otherwise the trivial estimator $\wh{f} = 0$ achieves smaller error than the parametric rate $n^{-1}$, and the problem doesn't fit well within the nonparametric setup.)
-\\
 %\sbcomment{I don't think the second part of this comment is correct the way you've defined the Sobolev class.} \\
@@
 
-\sbcomment{Define a level-$\alpha$ test.
+%\sbcomment{Define a level-$\alpha$ test.
 You use it in~\eqref{eqn:type_I_error} for example.
@@
 Later on you seem to use $b$ for Type-I error.
-Maybe use $b$ for Type-II and $\alpha$ for Type-I.} The minimax critical radius is the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{F}_{\epsilon} := \mc{F} \cap \{f: \|f\|_{\Leb^2(\Xset)} \geq \epsilon\}$:
+Maybe use $b$ for Type-II and $\alpha$ for Type-I.}
+
+The Type I error of a test $\phi$ is $\mathbb{E}_0[\phi]$, and if $\mathbb{E}_0[\phi] \leq \alpha$ for a given $\alpha \in (0,1)$ we refer to $\phi$ as a level-$\alpha$ test.
+The worst-case risk of $\phi$ over $\mc{F}$ is
 \begin{equation*}
-\epsilon(\mc{F}) := \inf\biggl\{\epsilon > 0: \inf_{\phi} \sup_{f_0 \in \mc{F}_{\epsilon}} \Ebb_{f_0}[1 - \phi] \leq \alpha \biggr\},
+R_n(\phi, \mc{F}, \epsilon) := \sup\Bigl\{\mathbb{E}_{f_0}[1 - \phi]: f_0 \in \mc{F}, \|f_0\|_{\Leb^2(\mc{X})} > \epsilon\Bigr\},
 \end{equation*}
-where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Of course, the minimax critical radius $\epsilon$ depends on $\alpha$.
-However, we adopt the typical convention of treating $\alpha$ as a small but fixed positive constant; hence it will not affect the testing error rates, and we suppress it notationally.} See \citet{ingster82,ingster87,ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
+and for a given constant $b \geq 1$, the minimax critical radius is the smallest value of $\epsilon$ such that some level-$\alpha$ test has worst-case risk of at most $1/b$. Formally,
+\begin{equation*}
+\epsilon(\mc{F}) := \inf\Bigl\{\epsilon > 0: \inf_{\phi} R_n(\phi,\mc{F},\epsilon) \leq 1/b \Bigr\},
+\end{equation*} 
+where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Clearly, the minimax critical radius $\epsilon$ depends on $\alpha$ and $b$.
+However, we adopt the typical convention of treating $\alpha$ and $b$ as small but fixed positive constants; hence it will not affect the testing error rates, and we suppress it notationally.} See \citet{ingster82,ingster87,ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
 
@@
 \end{equation*}
-where we recall $\lambda_k$ is the $k$th smallest eigenvalue of \smash{$\Lap_{n,r}$}. The Laplacian smoothing test is then simply
+where we recall $\lambda_k$ is the $k$th smallest eigenvalue of \smash{$L$}. The Laplacian smoothing test is then simply
 \begin{equation*}
@@
 \end{enumerate}
-We will also require that the radius of the Sobolev class not be too large. \\
-%\sbcomment{What does the class have to do with graphs?}  \\
-%\agcomment{Resolved.} \\
-Precisely, we will require $M \leq M_{\max}(d)$, where we define
+We will also require that the radius of the Sobolev class not be too large. Precisely, we will require $M \leq M_{\max}(d)$, where we define
 \begin{equation*}
@@
 \begin{aligned}
-\lim \frac{1}{n^2 r^{d + 2}} f^\top \Lap_{n,r} f & = \int_{\Xset} f(x) \cdot \Delta_Pf(x) p(x) \,dx \\
+\lim \frac{1}{n^2 r^{d + 2}} f^\top L f & = \int_{\Xset} f(x) \cdot \Delta_Pf(x) p(x) \,dx \\
 & = \int_{\Xset} \|\nabla f(x)\|_2^2 p^2(x) \,dx.
@@
 \begin{equation*}
-\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \underbrace{\vphantom{\sum_{k=1}^n}\frac{2\rho}{n} \bigl(f_0^\top \Lap_{n,r} f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{10}{n} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2}}_{\textrm{variance}},
+\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \underbrace{\vphantom{\sum_{k=1}^n}\frac{2\rho}{n} \bigl(f_0^\top L f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{10}{n} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2}}_{\textrm{variance}},
 \end{equation*}
@@
 \begin{equation*}
-\bigl\|f_0\bigr\|_n^2 \geq \underbrace{\vphantom{\sqrt{\sum_{k=1}^n}}\frac{2\rho}{n} \bigl(f_0^\top \Lap_{n,r} f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{4b}{n} \sqrt{\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}}}_{\textrm{variance}}.
+\bigl\|f_0\bigr\|_n^2 \geq \underbrace{\vphantom{\sqrt{\sum_{k=1}^n}}\frac{2\rho}{n} \bigl(f_0^\top L f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{4b}{n} \sqrt{\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}}}_{\textrm{variance}}.
 \end{equation*}
@@
 	\label{eqn:graph_sobolev_seminorm}
-	f^\top \Lap_{n,r} f \leq \frac{C_2}{\delta} n^2 r^{d + 2} |f|_{H^1(\Xset)}^2.
+	f^\top L f \leq \frac{C_2}{\delta} n^2 r^{d + 2} |f|_{H^1(\Xset)}^2.
 	\end{equation}
@@
 \begin{equation*}
-H^1(G_{n,r},M) := \{f: f^\top \Lap_{n,r} f \leq M^2\}
+H^1(G_{n,r},M) := \{f: f^\top L f \leq M^2\}
 \end{equation*}
@@
 In practice, it is more common to use a $k$-nearest-neighbor (kNN) graph than a neighborhood graph, due to the guaranteed connectivity and sparsity of the former; we suspect that by building on the work of \citet{calder2019}, one can show that our main results all hold under the kNN graph as well.
-In another direction, one can also generalize Laplacian smoothing by replacing the penalty \smash{$f^\top \Lap_{n,r} f$} with \smash{$f^\top \Lap_{n,r}^{s} f$}, for an integer $s > 1$.
+In another direction, one can also generalize Laplacian smoothing by replacing the penalty \smash{$f^\top L f$} with \smash{$f^\top L^{s} f$}, for an integer $s > 1$.
 The hope is that this would then achieve minimax optimal rates over the higher-order Sobolev class $H^s(\Xset)$.

### 2021-03-08 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
Nitpicky edits.

@@
 
-Testing $f_0=0$ is an easier problem than estimating $f_0$, and hence the minimax testing critical radius over $H^1(\Xset,M)$ is smaller than the minimax estimation rate, for $1 \leq d < 4$ (see \citet{ingster2009}):
+Testing $f_0 = 0$ is an easier problem than estimating $f_0$, and hence the minimax testing critical radius over $H^1(\Xset,M)$ is smaller than the minimax estimation rate, for $1 \leq d < 4$ (see \citet{ingster2009}):
 \begin{equation}
@@
 Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume the neighborhood graph \smash{$G_{n,r}$} is computed with a kernel $K$ satisfying \ref{asmp:kernel}.
-There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_estimation}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing} with \smash{$\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$},
-	with probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c (M^2n)^{d/(2 + d)})$, satisfies
+There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_estimation}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing} with \smash{$\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$} satisfies
 	\begin{equation*}
@@
 	\end{equation*}
+	with probability at least $1 - \delta -  C_1n\exp(-c_1nr^d) - \exp(-c (M^2n)^{d/(2 + d)})$.
 	% RJT: Alden, please verify that what I wrote is correct (I reorganized the way we explained this, added for $n \geq N$ ...).
@@
 \end{equation*}
-and likewise that $\wh{\varphi}$ has high power whenever
+and likewise that $\wh{\varphi}$ has small risk whenever
 \begin{equation*}
@@
 Lemma~\ref{lem:neighborhood_eigenvalue} leads to a sufficiently tight upper bound on the variance term whenever the radius $r$ is sufficiently small; precisely, when $r$ is upper bounded as in \ref{asmp:ls_kernel_radius_estimation} for estimation, or \ref{asmp:ls_kernel_radius_testing} for testing.
-The parameter $\rho$ is then chosen to minimize the sum of these upper bounds on bias and variance, as usual, and some straightforward calculations give Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing}.\\
+The parameter $\rho$ is then chosen to minimize the sum of these upper bounds on bias and variance, as usual, and some straightforward calculations give Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing}.
 %\sbcomment{Should probably say something about the rest of the proof, i.e. is it straightforward to put these together to get the two Theorems?} \\

### 2021-03-09 manuscripts/laplacian_smoothing/arXiv/appendix.tex
Completed restatement of testing results. Changed appendix to match new statements.

@@
 \begin{dcases*}
-\lim_{\varepsilon \to 0} \frac{1}{\nu(B(x,\varepsilon))} \int_{B(x,\varepsilon)} f(x) dx,&~~\textrm{if the limit exists,} \\
+\lim_{\varepsilon \to 0} \frac{1}{\nu(B(x,\varepsilon))} \int_{B(x,\varepsilon)} f(z) dz,&~~\textrm{if the limit exists,} \\
 0,&~~\textrm{otherwise.}
@@
 	\label{lem:linear_smoother_fixed_graph_testing}
-	Let $\wc{T} = {\bf Y}^{\top} S^2 {\bf Y}$ for a square, symmetric matrix $S \in \Reals^{n \times n}$. Suppose $S$ is a contraction. Define the threshold $\wc{t}_b$ to be 
+	Let $\wc{T} = {\bf Y}^{\top} S^2 {\bf Y}$ for a square, symmetric matrix $S \in \Reals^{n \times n}$. Suppose $S$ is a contraction. Define the threshold $\wc{t}_{\alpha}$ to be 
 	\begin{equation}
 	\label{eqn:linear_smoother_fixed_graph_threshold}
-	\wc{t}_{b} := \norm{\lambdavec(S)}_2^2 + 2b \norm{\lambdavec(S)}_4^2.
+	\wc{t}_{\alpha} := \norm{\lambdavec(S)}_2^2 + \sqrt{\frac{2}{\alpha}} \norm{\lambdavec(S)}_4^2.
 	\end{equation}
@@
 		\label{eqn:linear_smoother_fixed_graph_testing_typeI}
-		\Pbb_0\bigl(\wc{T} > \wc{t}_b\bigr) \leq \frac{1}{b^2}.
+		\Pbb_0\bigl(\wc{T} > \wc{t}_{\alpha}\bigr) \leq \alpha.
 		\end{equation}
@@
 		\label{eqn:linear_smoother_fixed_graph_testing_critical_radius}
-		f_0^{\top} S^2 f_0 \geq 4b \norm{\lambdavec(S)}_4^2,
+		f_0^{\top} S^2 f_0 \geq \Bigl(2\sqrt{\frac{2}{\alpha}} + 2b\Bigr) \cdot \norm{\lambdavec(S)}_4^2,
 		\end{equation}
@@
 		\label{eqn:linear_smoother_fixed_graph_testing_typeII}
-		\Pbb_{f_0}\bigl(\wc{T} \leq \wc{t}_b\bigr) \leq \frac{4}{b^2} + \frac{8}{b \norm{\lambdavec(S)}_4^{2}}.
+		\Pbb_{f_0}\bigl(\wc{T} \leq \wc{t}_{\alpha}\bigr) \leq \frac{1}{b^2} + \frac{16}{b \norm{\lambdavec(S)}_4^{2}}.
 		\end{equation}
@@
 We now establish the upper bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeII} on Type II error.
-From assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}, we see that $f_0^{\top} S^2 f_0^{\top} - \wc{t}_b \leq 0$.
+From assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}, we see that $f_0^{\top} S^2 f_0^{\top} - \wc{t}_{\alpha} \leq 0$.
 As a result,
 \begin{align*}
-\Pbb_{f_0}\Bigl(\wc{T} \leq \wc{t}_b\Bigr) & = \Pbb_{f_0}\Bigl(\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr] \leq \wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr) \\ 
-& \leq \Pbb_{f_0}\Bigl(\Bigl|\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr| \geq \Bigl|\wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr|\Bigr) \\ 
-& \leq \frac{\Var_{f_0}\bigl[\wc{T}\bigr]}{\bigl(\wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\bigr)^2},
+\Pbb_{f_0}\Bigl(\wc{T} \leq \wc{t}_{\alpha}\Bigr) & = \Pbb_{f_0}\Bigl(\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr] \leq \wc{t}_{\alpha} - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr) \\ 
+& \leq \Pbb_{f_0}\Bigl(\Bigl|\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr| \geq \Bigl|\wc{t}_{\alpha} - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr|\Bigr) \\ 
+& \leq \frac{\Var_{f_0}\bigl[\wc{T}\bigr]}{\bigl(\wc{t}_{\alpha} - \Ebb_{f_0}\bigl[\wc{T}\bigr]\bigr)^2},
 \end{align*}
 where the last line follows from Chebyshev's inequality.
-Plugging in the expressions~\eqref{pf:linear_smoother_fixed_graph_testing_mean} and~\eqref{pf:linear_smoother_fixed_graph_testing_var} for the mean and variance of $\wc{T}$, as well as the definition of $\wc{t}_b$ in~\eqref{eqn:linear_smoother_fixed_graph_threshold}, we obtain that
+Plugging in the expressions~\eqref{pf:linear_smoother_fixed_graph_testing_mean} and~\eqref{pf:linear_smoother_fixed_graph_testing_var} for the mean and variance of $\wc{T}$, as well as the definition of $\wc{t}_{\alpha}$ in~\eqref{eqn:linear_smoother_fixed_graph_threshold}, we obtain that
 \begin{equation}
 \label{pf:linear_smoother_fixed_graph_testing2}
-\Pbb_{f_0}\Bigl(\wc{T} \leq \wc{t}_b\Bigr) \leq \frac{4\|\lambdavec(S)\|_4^4}{\bigl(f_0^{\top}S^2f_0 - 2b \norm{\lambdavec(S)}_4^2\bigr)^2} + \frac{8f_0^{\top}S^4f_0 }{\bigl(f_0^{\top}S^2f_0 - 2b \norm{\lambdavec(S)}_4^2\bigr)^2}.
-\end{equation}
-We now use the assumed lower bound $f_0^{\top} S^2 f_0 \geq 4b \norm{\lambdavec(S)}_4^2$ to separately upper bound each of the two terms on the right hand side of~\eqref{pf:linear_smoother_fixed_graph_testing2}.
+\Pbb_{f_0}\Bigl(\wc{T} \leq \wc{t}_{\alpha}\Bigr) \leq \frac{4\|\lambdavec(S)\|_4^4}{\bigl(f_0^{\top}S^2f_0 - \sqrt{2/\alpha} \norm{\lambdavec(S)}_4^2\bigr)^2} + \frac{8f_0^{\top}S^4f_0 }{\bigl(f_0^{\top}S^2f_0 - \sqrt{2/\alpha} \norm{\lambdavec(S)}_4^2\bigr)^2}.
+\end{equation}
+We now use the assumed lower bound $f_0^{\top} S^2 f_0 \geq (2\sqrt{2/\alpha} + 2b) \norm{\lambdavec(S)}_4^2$ to separately upper bound each of the two terms on the right hand side of~\eqref{pf:linear_smoother_fixed_graph_testing2}.
 It follows immediately that
@@
 \label{pf:linear_smoother_fixed_graph_testing2.5}
-\frac{4 \|\lambdavec(S)\|_4^4}{\bigl(f_0^{\top}S^2f_0 - 2b \norm{\lambdavec(S)}_4^2\bigr)^2} \leq \frac{1}{b^2},
+\frac{4 \|\lambdavec(S)\|_4^4}{\bigl(f_0^{\top}S^2f_0 - \sqrt{2/\alpha} \norm{\lambdavec(S)}_4^2\bigr)^2} \leq \frac{1}{b^2},
 \end{equation}
@@
 \label{pf:linear_smoother_fixed_graph_testing3}
-\frac{8f_0^{\top}S^4f_0 }{\bigl(f_0^{\top}S^2f_0 - 2b \norm{\lambdavec(S)}_4^2\bigr)^2} \leq \frac{32 f_0^{\top}S^4f_0 }{\bigl(f_0^{\top}S^2f_0\bigr)^2} \leq \frac{8}{b \|\lambdavec(S)\|_4^2} \frac{f_0^{\top}S^4f_0 }{f_0^{\top}S^2f_0} \leq \frac{8}{b \|\lambdavec(S)\|_4^2},
+\frac{8f_0^{\top}S^4f_0 }{\bigl(f_0^{\top}S^2f_0 - \sqrt{2/\alpha} \norm{\lambdavec(S)}_4^2\bigr)^2} \leq \frac{32 f_0^{\top}S^4f_0 }{\bigl(f_0^{\top}S^2f_0\bigr)^2} \leq \frac{16}{b \|\lambdavec(S)\|_4^2} \frac{f_0^{\top}S^4f_0 }{f_0^{\top}S^2f_0} \leq \frac{16}{b \|\lambdavec(S)\|_4^2},
 \end{equation}
@@
 \begin{equation*}
-\wh{t}_b := \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k + 1\bigr)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k + 1\bigr)^4}}.
+\wh{t}_{\alpha} := \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k + 1\bigr)^2} + \frac{1}{n}\sqrt{\frac{2}{\alpha}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k + 1\bigr)^4}}.
 \end{equation*}
@@
 		\label{eqn:ls_fixed_graph_testing_typeI}
-		\Pbb_0\Bigl(\wh{T} > \wh{t}_b\Bigr) \leq \frac{1}{b^2}.
+		\Pbb_0\Bigl(\wh{T} > \wh{t}_{\alpha}\Bigr) \leq \alpha.
 		\end{equation}
@@
 		\label{eqn:ls_fixed_graph_testing_critical_radius}
-		\frac{1}{n}\norm{f_0}_2^2 \geq \frac{2 \rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{4b}{n} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{1/2},
+		\frac{1}{n}\norm{f_0}_2^2 \geq \frac{2 \rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{2\sqrt{2/\alpha} + 2b}{n} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{1/2},
 		\end{equation}
@@
 		\label{eqn:ls_fixed_graph_testing_typeII}
-		\Pbb_{f_0}\Bigl(\wh{T}(G) \leq \wh{t}_b\Bigr) \leq \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2}.
+		\Pbb_{f_0}\Bigl(\wh{T}(G) \leq \wh{t}_{\alpha}\Bigr) \leq \frac{1}{b^2} + \frac{16}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2}.
 		\end{equation}
@@
 \end{equation*} 
-In the above, we have written $\Lap^{-1/2}$ for the square root of the pseudoinverse of $\Lap$, and the maximum is over all indices $k$ such that $\lambda_k > 0$.
+In the above, we have written $\Lap^{-1/2}$ for the square root of the pseudoinverse of $\Lap$, the maximum is over all indices $k$ such that $\lambda_k > 0$, and the last inequality follows from the basic algebraic identity $1 - 1/(1 + x)^2 \leq 2 x$ for any $x > 0$.
 The claim of the Lemma then follows from Lemma~\ref{lem:linear_smoother_fixed_graph_estimation}.
@@
 f_0^{\top} \wh{S}^2 f_0 & = \bigl\|f_0\bigr\|_2^2 - f_0^{\top}(I - \wh{S}^2)f_0 \\
-& \geq 2 \rho \bigl(f_0^{\top} \Lap f_0\bigr) - f_0^{\top}\bigl(I - \wh{S}^2\bigr)f_0 + 4b \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2}.
+& \geq 2 \rho \bigl(f_0^{\top} \Lap f_0\bigr) - f_0^{\top}\bigl(I - \wh{S}^2\bigr)f_0 +  \Bigl(2\sqrt{\frac{2}{\alpha}} + 2b\Bigr) \cdot \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2}.
 \end{align*}
@@
 \end{equation*}
----where in the above the maximum is over all indices $k$ such that $\lambda_k > 0$, and the last inequality follows from the basic algebraic identity $1 - 1/(1 + x)^2 \leq 2 x$ for any $x > 0$ \sbcomment{You've used this before, so maybe move this comment earlier?}---we deduce that
-\begin{equation*}
-f_0^{\top} \wh{S}^2 f_0 \geq 4b \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{-1/2}.
+---where in the above the maximum is over all indices $k$ such that $\lambda_k > 0$---we deduce that
+\begin{equation*}
+f_0^{\top} \wh{S}^2 f_0 \geq  \Bigl(2\sqrt{\frac{2}{\alpha}} + 2b\Bigr) \cdot \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{-1/2}.
 \end{equation*} 
@@
 First, we upper bound the right hand side of this inequality.
-Set $\wt{\rho} = M^{-8/(4 + d)}n^{-4/(4 + d)}$ as prescribed by Theorem~\ref{thm:laplacian_smoothing_testing}, it follows from \eqref{eqn:graph_sobolev_seminorm} and \eqref{eqn:neighborhood_eigenvalue} that
-\begin{align*}
-\frac{2 \rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{4b}{n} \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{1/2} & \leq C_2b\wt{\rho}M^2 + \frac{4b}{n}\Biggl[1 + \frac{1}{c_3^2}\biggl(\sum_{k = 2}^{n} \frac{1}{(\wt{\rho}k^{2/d} + 1)^4} \biggr)^{1/2} + \frac{r^4n^{1/2}}{c_3^2\wt{\rho}^2}\Biggr] \\
-& \leq C_2b\wt{\rho}M^2 + \frac{4b}{n}\biggl(1 + \frac{\sqrt{2}}{c_3^{2}}\wt{\rho}^{-d/4} + \frac{r^4n^{1/2}}{c_3^2\wt{\rho}^2}\biggr) \\ 
-& \leq \Bigl(C_2 + 4 + \frac{4\sqrt{2}}{c_3^{2}} + \frac{4}{c_3^2}\Bigr)bM^{2d/(4 + d)} n^{-4/(4 + d)}.
+Setting $\wt{\rho} = M^{-8/(4 + d)}n^{-4/(4 + d)}$ as prescribed by Theorem~\ref{thm:laplacian_smoothing_testing}, it follows from \eqref{eqn:graph_sobolev_seminorm} and \eqref{eqn:neighborhood_eigenvalue} that
+\begin{align*}
+\frac{2 \rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{2\sqrt{2/\alpha} + 2b}{n} \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{1/2} & \leq C_2b\wt{\rho}M^2 + \frac{2\sqrt{2/\alpha} + 2b}{n}\Biggl[1 + \frac{1}{c_3^2}\biggl(\sum_{k = 2}^{n} \frac{1}{(\wt{\rho}k^{2/d} + 1)^4} \biggr)^{1/2} + \frac{r^4n^{1/2}}{c_3^2\wt{\rho}^2}\Biggr] \\
+& \leq C_2b\wt{\rho}M^2 + \frac{2\sqrt{2/\alpha} + 2b}{n}\biggl(1 + \frac{\sqrt{2}}{c_3^{2}}\wt{\rho}^{-d/4} + \frac{r^4n^{1/2}}{c_3^2\wt{\rho}^2}\biggr) \\ 
+& \leq \Bigl(C_2 + 2 + \frac{2\sqrt{2}}{c_3^{2}} + \frac{2}{c_3^2}\Bigr)\cdot \Bigl(\sqrt{\frac{2}{\alpha}} + b\Bigr) \cdot M^{2d/(4 + d)} n^{-4/(4 + d)}.
 \end{align*}
@@
 Recall that by assumption, $M \leq M_{\max}(d)$.
-Therefore, taking $C \geq C_6$ in \eqref{eqn:laplacian_smoothing_testing}, implies that the lower bound on $\|f\|_{\Leb^2(\mc{X})}$ in \eqref{eqn:empirical_norm_sobolev_1} is satisfied.
+Therefore, taking $C \geq C_6$ in \eqref{eqn:laplacian_smoothing_testing} implies that the lower bound on $\|f\|_{\Leb^2(\mc{X})}$ in \eqref{eqn:empirical_norm_sobolev_1} is satisfied.
 As a result, it follows from \eqref{eqn:empirical_norm_sobolev} that
 \begin{equation*}
-\|f\|_n^2 \geq \frac{\mathbb{E}[\|f\|_n^2]}{b} \geq \frac{p_{\min}}{b} \|f\|_{\Leb^2(\Xset)}^2 \geq CbM^{2d/(4 + d)}n^{-4/(4 + d)},
+\|f\|_n^2 \geq \frac{\mathbb{E}[\|f\|_n^2]}{b} \geq \frac{p_{\min}}{b} \|f\|_{\Leb^2(\Xset)}^2 \geq C\Bigl(\sqrt{\frac{1}{\alpha}} + b\Bigr)M^{2d/(4 + d)}n^{-4/(4 + d)},
 \end{equation*}
 with probability at least $1 - 5/b$.
-Taking $C \geq C_2 + 4 + (4\sqrt{2})/c_3^{2} + 4/c_3^2$ in \eqref{eqn:laplacian_smoothing_testing} thus implies \eqref{eqn:ls_fixed_graph_testing_critical_radius}, and we may therefore use Lemma~\ref{lem:ls_fixed_graph_testing} to upper bound the type II error the Laplacian smoothing test $\wh{\varphi}$.
+Taking $C \geq C_2 + 2 + (2\sqrt{2})/c_3^{2} + 2/c_3^2$ in \eqref{eqn:laplacian_smoothing_testing} thus implies \eqref{eqn:ls_fixed_graph_testing_critical_radius}, and we may therefore use Lemma~\ref{lem:ls_fixed_graph_testing} to upper bound the type II error the Laplacian smoothing test $\wh{\varphi}$.
 Observe that by \eqref{eqn:neighborhood_eigenvalue} and the lower bound in Lemma~\ref{lem:variance_term_testing},
@@
 \begin{align*}
-\Pbb_{f_0}\bigl(\wh{T} \leq \wh{t}_b\bigr) & \leq \frac{6}{b} + \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2} + C_1n\exp(-c_1nr^d)\\
-& \leq \frac{10}{b} + \frac{32\sqrt{2}}{b} C_3^{2}\wt{\rho}^{d/4} + C_1n\exp(-c_1nr^d),
+\Pbb_{f_0}\bigl(\wh{T} \leq \wh{t}_{\alpha}\bigr) & \leq \frac{6}{b} + \frac{1}{b^2} + \frac{16}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2} + C_1n\exp(-c_1nr^d)\\
+& \leq \frac{7}{b} + \frac{64\sqrt{2}}{b} C_3^{2}\wt{\rho}^{d/4} + C_1n\exp(-c_1nr^d),
 \end{align*}
@@
 \paragraph{Proof of \eqref{eqn:laplacian_smoothing_testing_low_smoothness}.}
-When $\rho = 0$, the Laplacian smoother $\wh{f} = \mathbf{Y}$, the test statistic $\wh{T} = \frac{1}{n}\|\mathbf{Y}\|_2^2$, and the threshold $\wh{t}_b = 1 + 2bn^{-1/2}$.
+When $\rho = 0$, the Laplacian smoother $\wh{f} = \mathbf{Y}$, the test statistic $\wh{T} = \frac{1}{n}\|\mathbf{Y}\|_2^2$, and the threshold $\wh{t}_{\alpha} = 1 + 2bn^{-1/2}$.
 The expectation of $\wh{T}$ is
@@
 \end{equation*}
-so that $E[\wh{T}] - \wh{t}_b \geq E[f_0^2(X)]/2$. Hence, by Chebyshev's inequality
-\begin{align*}
-\mathbb{P}_{f_0}\Bigl(\wh{T} \leq \wh{t}_b \Bigr) & \leq 4 \frac{\Var_{f_0}\bigl[\wh{T}\bigr]}{\mathbb{E}[f_0^2(X)]^2} \\
+so that $E[\wh{T}] - \wh{t}_{\alpha} \geq E[f_0^2(X)]/2$. Hence, by Chebyshev's inequality
+\begin{align*}
+\mathbb{P}_{f_0}\Bigl(\wh{T} \leq \wh{t}_{\alpha} \Bigr) & \leq 4 \frac{\Var_{f_0}\bigl[\wh{T}\bigr]}{\mathbb{E}[f_0^2(X)]^2} \\
 & \leq \frac{4}{n} \cdot \frac{ 3 + p_{\max}\bigl(M^4 + \|f_0\|_{\Leb^2(\Xset)}^2 \bigr)}{p_{\min}^2 \|f_0\|_{\Leb^2(\Xset)}^4} \\

### 2021-03-09 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
Completed restatement of testing results. Changed appendix to match new statements.

@@
 	\begin{equation*}
-	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(2 + d)} n^{-2/(2 + d)}.
+	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(2 + d)} n^{-2/(2 + d)},
 	\end{equation*}
@@
 
-For $b \geq 1$, define a threshold \smash{$\wh{t}_b$} as
+For a given $0 < \alpha < 1$, define a threshold \smash{$\wh{t}_{\alpha}$} as
 \begin{equation*}
-\wh{t}_{b} = \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}},
+\wh{t}_{\alpha} = \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2} + \frac{1}{n}\sqrt{\frac{2}{\alpha} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}},
 \end{equation*}
@@
 \begin{equation*}
-\wh{\varphi} = \1\bigl\{\wh{T} \leq \wh{t}_b\bigr\}.
+\wh{\varphi} = \1\bigl\{\wh{T} \leq \wh{t}_{\alpha}\bigr\}.
 \end{equation*} 
 %(Clearly \smash{$\wh{\varphi}$} depends on $b$, but we suppress this notationally.) 
-Here $b \geq 1$ is a tuning parameter selected by the user, with the choice driven by the tolerated Type I versus Type II error.
-In fact, the threshold \smash{$\wh{t}_b$} is precisely the right choice to control the Type I error of \smash{$\wh{\varphi}$}: we show in Appendix~\ref{sec:fixed_graph_error_bounds} that
-\begin{equation}
-\label{eqn:type_I_error}
-\Ebb_0\bigl[\wh{\varphi}\bigr] \leq \frac{1}{b^2}.
-\end{equation}
-The next theorem shows that this test also has small Type II error, uniformly over all $f_0$ that are separated from $0$ by at least the critical radius given in \eqref{eqn:sobolev_space_testing_critical_radius}.
+We show in Appendix~\ref{sec:fixed_graph_error_bounds} that $\wh{f}$ is a level-$\alpha$ test.
+In the next theorem, we upper bound the worst-case risk $R_n(\wh{\varphi},H^1(\mc{X},M),\epsilon)$ of $\wh{\varphi}$, whenever $\epsilon$ is at least (a constant times) the critical radius given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
 For this to hold, we will require a tighter range of scalings for the graph radius $r$.
@@
 Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume \smash{$G_{n,r}$} is computed with a kernel $K$ satisfying \ref{asmp:kernel}.
-There are constants $N,C,c>0$ such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_testing}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$}, satisfies the following: for all $b \geq 1$, and all $f_0$ such that
+There are constants $N,C,C_1,c_1>0$ such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_testing}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$}, satisfies the following: for any $b \geq 1$, if
 	\begin{equation}
 	\label{eqn:laplacian_smoothing_testing}
-	\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2d/(4 + d)} n^{-4/(4 + d)},
+	\epsilon^2 \geq C M^{2d/(4 + d)} n^{-4/(4 + d)}\Bigl(b^2 + b\sqrt{\frac{1}{\alpha}}\Bigr) ,
 	\end{equation} 
-	the Type II error is upper bounded by
-	\begin{equation*}
-	\Ebb_{f_0}\bigl[1 - \wh{\varphi} \bigr] \leq C\biggl(\frac{1}{b}\Bigl[1 + n^{-d/(4 + d)}M^{-2d/(4 + d)}\Bigr] + n\exp\bigl(-cnr^d\bigr)\biggr).
-	\end{equation*}
+	then the worst-case risk is upper bounded, $R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + C_1n\exp(-c_1nr^d)\bigr)$.
 \end{theorem}
-\sbcomment{Personally, I wouldn't emphasize the Type II error so much in the Theorem (it is not tight and it's not an interesting quantity anyway), and write \eqref{eqn:laplacian_smoothing_testing} more directly as an upper bound on the critical radius (which is the interesting quantity).
+%\sbcomment{Personally, I wouldn't emphasize the Type II error so much in the Theorem (it is not tight and it's not an interesting quantity anyway), and write \eqref{eqn:laplacian_smoothing_testing} more directly as an upper bound on the critical radius (which is the interesting quantity).
 Otherwise you've introduced that terminology for no reason.}
@@
 	As in Theorem \ref{thm:laplacian_smoothing_testing}, but where $\Xset \subseteq \Rd$ is a manifold with intrinsic dimension $m < 4$, $M \leq M_{\max}(m)$, and the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}.
-	There are constants $N,C,c>0$ such that for any $n \geq N$, and any $r$ as in \ref{asmp:ls_kernel_radius_testing_manifold}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$}, satisfies the following: for all $b \geq 1$, and $f_0$ such that
+	There are constants $N,C,C_1,c_1>0$ such that for any $n \geq N$, and any $r$ as in \ref{asmp:ls_kernel_radius_testing_manifold}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$}, satisfies the following: for any $b \geq 1$, if
 	\begin{equation}
 	\label{eqn:laplacian_smoothing_testing_manifold}
-	\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2m/(4 + m)} n^{-4/(4 + m)},
+	\epsilon^2 \geq C M^{2d/(4 + d)} n^{-4/(4 + d)}\Bigl(b^2 + b\sqrt{\frac{1}{\alpha}}\Bigr) ,
 	\end{equation} 
-	the Type II error is upper bounded by
-	\begin{equation*}
-	\Ebb_{f_0}\bigl[1 - \wh{\varphi} \bigr] \leq C\biggl(\frac{1}{b}\Bigl[1 + n^{-m/(4 + m)}M^{-2m/(4 + m)}\Bigr] + n\exp\bigl(-cnr^m\bigr)\biggr).
-	\end{equation*}
+	then the worst-case risk is upper bounded, $R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + C_1n\exp(-c_1nr^d)\bigr)$.
 	% RJT: Alden, please verify that what I wrote is correct.
@@
 \end{theorem}
-\sbcomment{Same comment as before on what to emphasize.}
+% \sbcomment{Same comment as before on what to emphasize.}
 
-The proof of Theorems \ref{thm:laplacian_smoothing_estimation_manifold} and \ref{thm:laplacian_smoothing_testing_manifold} proceeds in a similar manner to that of Theorems \ref{thm:laplacian_smoothing_estimation1} and \ref{thm:laplacian_smoothing_testing}.
+The proofs of Theorems \ref{thm:laplacian_smoothing_estimation_manifold} and \ref{thm:laplacian_smoothing_testing_manifold} proceed in a similar manner to that of Theorems \ref{thm:laplacian_smoothing_estimation1} and \ref{thm:laplacian_smoothing_testing}.
 The key difference is that in the manifold setting, the equations \eqref{eqn:graph_sobolev_seminorm} and \eqref{eqn:neighborhood_eigenvalue} used to upper bound bias and variance will hold with $d$ replaced by $m$.

### 2021-03-09 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
Fixed typo in last commit.

@@
 Under such an assumption, it is known \citep{bickel2007,ariascastro2018} that the optimal rates over $H^1(\Xset)$ are now $n^{-2/(2 + m)}$ (for estimation) and $n^{-4/(4 + m)}$ (for testing), which are much faster than the full-dimensional error rates when $m \ll d$.
-\sbcomment{I might point out that the Arias-Castro result is for density testing over Holder spaces, and that technically rates for regression testing on manifolds for Sobolev spaces are your contribution.}
+
+% \sbcomment{I might point out that the Arias-Castro result is for density testing over Holder spaces, and that technically rates for regression testing on manifolds for Sobolev spaces are your contribution.}
 
@@
 	\label{eqn:laplacian_smoothing_testing_manifold}
-	\epsilon^2 \geq C M^{2d/(4 + d)} n^{-4/(4 + d)}\Bigl(b^2 + b\sqrt{\frac{1}{\alpha}}\Bigr) ,
+	\epsilon^2 \geq C M^{2m/(4 + m)} n^{-4/(4 + m)}\Bigl(b^2 + b\sqrt{\frac{1}{\alpha}}\Bigr) ,
 	\end{equation} 
-	then the worst-case risk is upper bounded, $R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + C_1n\exp(-c_1nr^d)\bigr)$.
+	then the worst-case risk is upper bounded, $R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + C_1n\exp(-c_1nr^m)\bigr)$.
 	% RJT: Alden, please verify that what I wrote is correct.

### 2021-03-09 manuscripts/laplacian_smoothing/arXiv/appendix.tex
Finished responding to Siva's comments.

@@
 
-In this section, we prove Lemma~\ref{lem:graph_sobolev_seminorm}, which states an upper bound on $f^{\top} \Lap_{n,r} f$ that holds when $f$ is bounded in Sobolev norm.
+In this section, we prove Lemma~\ref{lem:graph_sobolev_seminorm}, which states an upper bound on $f^{\top} L f$ that holds when $f$ is bounded in Sobolev norm.
 We also establish stronger bounds in the case when $f$ has a bounded Lipschitz constant; this latter result justifies one of our remarks after Theorem~\ref{thm:laplacian_smoothing_estimation1}.
@@
 \begin{equation*}
-f^{\top} L_{n,r} f = g^{\top} L_{n,r} g \leq \frac{C_2}{\delta} n^2 r^{d + 2} |g|_{H^1(\mc{X})}^2 = \frac{C_2}{\delta} n^2 r^{d + 2} |f|_{H^1(\mc{X})}^2.
+f^{\top} L f = g^{\top} L g \leq \frac{C_2}{\delta} n^2 r^{d + 2} |g|_{H^1(\mc{X})}^2 = \frac{C_2}{\delta} n^2 r^{d + 2} |f|_{H^1(\mc{X})}^2.
 \end{equation*} 
@@
 \begin{equation*}
-\Ebb\Bigl[f^{\top} \Lap_{n,r} f\Bigr] \leq C n^2 r^{d + 2} \|f\|_{H^1(\Xset)}^2,
-\end{equation*}
-since the high-probability upper bound then follows immediately by Markov's inequality.
-(Recall that $\Lap_{n,r}$ is positive semi-definite, and therefore $f^{\top} \Lap_{n,r} f$ is a non-negative random variable).
+\Ebb\Bigl[f^{\top} L f\Bigr] \leq C n^2 r^{d + 2} \|f\|_{H^1(\Xset)}^2,
+\end{equation*}
+since the high-probability upper bound then follows immediately by Markov's inequality. (Recall that $L$ is positive semi-definite, and therefore $f^{\top} L f$ is a non-negative random variable).
 
@@
 \begin{equation*}
-f^{\top} \Lap_{n,r} f = \frac{1}{2}\sum_{i, j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 W_{ij},
+f^{\top} L f = \frac{1}{2}\sum_{i, j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 W_{ij},
 \end{equation*}
@@
 \label{pf:first_order_graph_sobolev_seminorm_1}
-\Ebb\Bigl[f^{\top} \Lap_{n,r} f\Bigr] = \frac{n(n - 1)}{2} \Ebb\biggl[\Bigl(f(X') - f(X)\Bigr)^2 K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr],
+\Ebb\Bigl[f^{\top} L f\Bigr] = \frac{n(n - 1)}{2} \Ebb\biggl[\Bigl(f(X') - f(X)\Bigr)^2 K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr],
 \end{equation}
@@
 Suppose $f$ satisfies $|f(x') - f(x)| \leq M \|x - x\|$ for all $x,x' \in \mc{X}$.
-Then we can strengthen the high probability bound in Lemma~\ref{lem:graph_sobolev_seminorm} from $1 - \delta$ to $1 - \delta^2/n$, at the cost of only a constant factor in the upper bound on $f^{\top} \Lap_{n,r} f$.
+Then we can strengthen the high probability bound in Lemma~\ref{lem:graph_sobolev_seminorm} from $1 - \delta$ to $1 - \delta^2/n$, at the cost of only a constant factor in the upper bound on $f^{\top} L f$.
 \begin{proposition}
@@
 	\begin{equation*}
-	f^{\top} \Lap_{n,r} f \leq \biggl(\frac{1}{\delta} + C_2\biggr) n^2 r^{d + 2} M^2.
+	f^{\top} L f \leq \biggl(\frac{1}{\delta} + C_2\biggr) n^2 r^{d + 2} M^2.
 	\end{equation*}
@@
 \paragraph{Proof of Proposition~\ref{prop:graph_sobolev_seminorm_lipschitz}.}
-We will prove Proposition~\ref{prop:graph_sobolev_seminorm_lipschitz} using Chebyshev's inequality, so the key step is to upper bound the variance of $f^{\top}\Lap_{n,r}f$.
-Putting $\varDelta_{ij} := K(\|X_i - X_j\|/r) \cdot (f(X_i) - f(X_j))^2$, we can write the variance of $f^{\top} \Lap_{n,r} f$ as a sum of covariances,
-\begin{equation*}
-\Var\bigl[f^{\top} \Lap_{n,r} f\bigr] = \frac{1}{4} \sum_{i, j = 1}^{n} \sum_{\ell,m = 1}^{n} \Cov\bigl[\varDelta_{ij},\varDelta_{\ell m}\bigr].
+We will prove Proposition~\ref{prop:graph_sobolev_seminorm_lipschitz} using Chebyshev's inequality, so the key step is to upper bound the variance of $f^{\top}Lf$.
+Putting $\varDelta_{ij} := K(\|X_i - X_j\|/r) \cdot (f(X_i) - f(X_j))^2$, we can write the variance of $f^{\top} L f$ as a sum of covariances,
+\begin{equation*}
+\Var\bigl[f^{\top} L f\bigr] = \frac{1}{4} \sum_{i, j = 1}^{n} \sum_{\ell,m = 1}^{n} \Cov\bigl[\varDelta_{ij},\varDelta_{\ell m}\bigr].
 \end{equation*}
@@
 \begin{equation*}
-\Var\bigl[f^{\top} \Lap_{n,r} f\bigr] \leq n^3 p_{\max}^3 M^4 r^{4 + 2d} + n^2 p_{\max}^2 M^4 r^{4 + d} K(0) \leq C M^4 n^3r^{4 + 2d},
+\Var\bigl[f^{\top} L f\bigr] \leq n^3 p_{\max}^3 M^4 r^{4 + 2d} + n^2 p_{\max}^2 M^4 r^{4 + d} K(0) \leq C M^4 n^3r^{4 + 2d},
 \end{equation*}
@@
 \begin{equation*}
-\Pbb\biggl(\Bigl|f^{\top} \Lap_{n,r} f - \Ebb\bigl[f^{\top} \Lap_{n,r} f\bigr]\Bigr| \geq \frac{M^2}{\delta}n^2 r^{d + 2}\biggr) \leq C\frac{\delta^2}{n},
-\end{equation*}
-and since we have already upper bounded $\Ebb\bigl[f^{\top} \Lap_{n,r} f\bigr] \leq C_2 M^2 n^2 r^{d + 2}$, the proposition follows.
+\Pbb\biggl(\Bigl|f^{\top} L f - \Ebb\bigl[f^{\top} L f\bigr]\Bigr| \geq \frac{M^2}{\delta}n^2 r^{d + 2}\biggr) \leq C\frac{\delta^2}{n},
+\end{equation*}
+and since we have already upper bounded $\Ebb\bigl[f^{\top} L f\bigr] \leq C_2 M^2 n^2 r^{d + 2}$, the proposition follows.
 
@@
 In this section, we prove Lemma~\ref{lem:neighborhood_eigenvalue}, following the lead of~\citet{burago2014,trillos2019,calder2019}, who establish similar results with respect to a manifold without boundary.
-To prove this lemma, in  Theorem~\ref{thm:neighborhood_eigenvalue} we give estimates on the difference between eigenvalues of the graph Laplacian $L_{n,r}$ and eigenvalues of the weighted Laplace-Beltrami operator $\Delta_P$.
+To prove this lemma, in  Theorem~\ref{thm:neighborhood_eigenvalue} we give estimates on the difference between eigenvalues of the graph Laplacian $L$ and eigenvalues of the weighted Laplace-Beltrami operator $\Delta_P$.
 We recall $\Delta_P$ is defined as
@@
 \end{equation*}
-To avoid confusion, in this section we write $\lambda_k(G_{n,r})$ for the $k$th smallest eigenvalue of the graph Laplacian matrix $\Lap_{n,r}$ and $\lambda_k(\Delta_P)$ for the $k$th smallest eigenvalue of $\Delta_P$ \footnote{Under the assumptions~\ref{asmp:domain} and~\ref{asmp:density}, the operator $\Delta_P$ has a discrete spectrum; see \citet{garciatrillos18} \sbcomment{Fix this reference, typos.} for more details.}.
+To avoid confusion, in this section we write $\lambda_k(G_{n,r})$ for the $k$th smallest eigenvalue of the graph Laplacian matrix $L$ and $\lambda_k(\Delta_P)$ for the $k$th smallest eigenvalue of $\Delta_P$ \footnote{Under the assumptions~\ref{asmp:domain} and~\ref{asmp:density}, the operator $\Delta_P$ has a discrete spectrum; see \citet{garciatrillos18} for more details.}.
 Some other notation: throughout this section, we will write $A, A_0, A_1,\ldots$ and $a,a_0,a_1,\ldots$ for constants which may depend on $\Xset$, $d$, $K$, and $p$, but do not depend on $n$; we keep track of all such constants explicitly in our proofs.
@@
 
-\agcomment{Siva, I erred on the side of giving too much detail here, figuring it was easier for you/me to pare back if needed.}
-\sbcomment{It is extremely insightful actually. I would consider including some of the second paragraph in the main paper too.}
-
 The bulk of the remainder of this section is devoted to the proof of Theorem~\ref{thm:neighborhood_eigenvalue}.
@@
 \begin{equation*}
-b_r(u) := \frac{1}{n^2 r^{d+ 2}}u^{\top} \Lap_{n,r} u 
+b_r(u) := \frac{1}{n^2 r^{d+ 2}}u^{\top} L u 
 \end{equation*}
@@
 The upper bound follows from essentially parallel reasoning.
-Recalling that $v_1,\ldots,v_k$ denote the first $k$ eigenvectors of $\Lap_{n,r}$, set $U := \mathrm{span}\{v_1,\ldots,v_k\}$, so that $nr^{d + 2} b_r(u) \leq \lambda_k(G_{n,r}) \|u\|_n^2$.
+Recalling that $v_1,\ldots,v_k$ denote the first $k$ eigenvectors of $L$, set $U := \mathrm{span}\{v_1,\ldots,v_k\}$, so that $nr^{d + 2} b_r(u) \leq \lambda_k(G_{n,r}) \|u\|_n^2$.
 By Proposition~\ref{prop:isometry}, Part (2), we have that for every $u \in U$,
@@
 	\begin{equation*}
-	f^{\top} \Lap_{n,r} f \leq \frac{C}{\delta} n^2 r^{m + 2} |f|_{H^1(\Xset)}^2.
+	f^{\top} L f \leq \frac{C}{\delta} n^2 r^{m + 2} |f|_{H^1(\Xset)}^2.
 	\end{equation*}
@@
 \paragraph{Proof of \eqref{eqn:laplacian_smoothing_testing_low_smoothness}.}
-When $\rho = 0$, the Laplacian smoother $\wh{f} = \mathbf{Y}$, the test statistic $\wh{T} = \frac{1}{n}\|\mathbf{Y}\|_2^2$, and the threshold $\wh{t}_{\alpha} = 1 + 2bn^{-1/2}$.
+When $\rho = 0$, the Laplacian smoother $\wh{f} = \mathbf{Y}$, the test statistic $\wh{T} = \frac{1}{n}\|\mathbf{Y}\|_2^2$, and the threshold $\wh{t}_{\alpha} = 1 + n^{-1/2}\sqrt{2/\alpha}$.
 The expectation of $\wh{T}$ is
@@
 \begin{equation*}
-\norm{f_0}_{\Leb^2(X)}^2 \geq \frac{4b}{p_{\min}} n^{-1/2},
+\norm{f_0}_{\Leb^2(X)}^2 \geq \frac{2\sqrt{2/\alpha} + 2b}{p_{\min}} n^{-1/2},
 \end{equation*}
@@
 & \leq \frac{4}{n} \cdot \frac{ 3 + p_{\max}\bigl(M^4 + \|f_0\|_{\Leb^2(\Xset)}^2 \bigr)}{p_{\min}^2 \|f_0\|_{\Leb^2(\Xset)}^4} \\
-& \leq \frac{1}{16b^2}\Bigl(3 + \frac{4bp_{\max}}{p_{\min}n^{1/2}} + p_{\max}M^4\Bigr).
+& \leq \frac{1}{b^2}\Bigl(3 + \frac{4bp_{\max}}{p_{\min}n^{1/2}} + p_{\max}M^4\Bigr).
 \end{align*}

### 2021-03-09 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
Finished responding to Siva's comments.

@@
 \section{Summary of Results}
+\label{sec:summary}
 
@@
 \end{equation}
-where the optimization is performed over all functions for which the criterion is well-defined and finite (continuous functions $f$ that have a weak derivative $\nabla f$ in $L^2(\Xset)$).
-Analogously, for testing, we could use:
+where the optimization is performed over all continuous functions $f$ that have a weak derivative $\nabla f$ in $L^2(\Xset)$. Analogously, for testing, we could use:
 \begin{equation}
@@
 By contrast, smoothing splines are optimal only when $d=1$.
-When $d>1$, thin-plate spline estimator \eqref{eqn:thin_plate_spline} is not even well-posed, in the following sense: for any $(X_1,Y_1),\ldots,(X_n,Y_n)$ and any $\delta>0$, there exists (e.g., \citet{green93} give a construction using ``bump'' functions) a differentiable function $f$ such that $f(X_i) = Y_i$, $i=1,\ldots,n$, and
+When $d>1$, the thin-plate spline estimator \eqref{eqn:thin_plate_spline} is not even well-posed, in the following sense: for any $(X_1,Y_1),\ldots,(X_n,Y_n)$ and any $\delta>0$, there exists (e.g., \citet{green93} give a construction using ``bump'' functions) a differentiable function $f$ such that $f(X_i) = Y_i$, $i=1,\ldots,n$, and
 \begin{equation*}
@@
 Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius of the Sobolev ball is also optimal.
+Rates for thin-plate splines with $d \geq 2$ assume the estimator $\wt{f}$ interpolates the responses, $\wt{f}(X_i) = Y_i$ for $i = 1,\ldots,n$; see the discussion in Section~\ref{sec:summary}.
 We use ``$1$'' to indicate inconsistency (error not converging to 0).
@@
 	\caption{Summary of testing rates over first-order Sobolev balls; black, red, and bold fonts are used as in Table~\ref{tbl:estimation_rates}.
+Rates for thin-plate splines with $d \geq 2$ assume the test statistic $\wt{T}$ is computed with respect to an estimator $\wt{f}$ that interpolates the responses, $\wt{f}(X_i) = Y_i$ for $i = 1,\ldots,n$.
 Rates for $d \geq 4$ assume that $f_0 \in \Leb^4(\Xset,M)$.
@@
 Lastly, we use $a \wedge b = \min\{a,b\}$.
-\\
 %\sbcomment{This is wrong, should check where it appears.} \\
@@
 We write the eigendecomposition of $L$ as \smash{$L = \sum_{k = 1}^{n} \lambda_k v_k v_k^\top$}, and we always assume, by convention, ordered eigenvalues $0 = \lambda_1 \leq \cdots \leq \lambda_n$, and unit-norm eigenvectors.
-\\
 %\sbcomment{Don't understand why you're switching between $L$ and $L_{n,r}$?} \\
@@
 \end{equation*}
-and for a given constant $b \geq 1$, the minimax critical radius is the smallest value of $\epsilon$ such that some level-$\alpha$ test has worst-case risk of at most $1/b$. Formally,
+and for a given constant $b \geq 1$, the minimax critical radius $\epsilon(\mc{F})$ is the smallest value of $\epsilon$ such that some level-$\alpha$ test has worst-case risk of at most $1/b$.
+Formally,
 \begin{equation*}
@@
 On the other hand, if we explicitly assume that $f_0 \in \Leb^4(\Xset,M)$, then \citet{guerre02} show that the testing problem is characterized by a dimension-free lower bound $\epsilon^{2}(\Leb^4(\Xset,M)) \gtrsim n^{-1/2}$.
-Moreover, by estimating \smash{$\wh{f}$}, setting $\rho = 0$, the subsequent test \smash{$\wh{\varphi}$} will achieve (up to constants) this lower bound.
-That is, for any $f_0 \in \Leb^4(\Xset,M)$ such that \smash{$\|f_0\|_{\Leb^2(\Xset)}^2 \geq C b^2n^{-1/2}$}, we have both
+Moreover, by setting $\rho = 0$ so that the resulting estimator $\wh{f}$ interpolates the responses $Y_1,\ldots,Y_n$, the subsequent test \smash{$\wh{\varphi}$} will achieve (up to constants) this lower bound.
+That is, for any $f_0 \in \Leb^4(\Xset,M)$ such that \smash{$\|f_0\|_{\Leb^2(\Xset)}^2 \geq C(b^2 + \sqrt{1/\alpha})n^{-1/2}$}, we have that \smash{$\Ebb_0[\wh{\varphi}] \leq \alpha$} and
 	\begin{equation}
 	\label{eqn:laplacian_smoothing_testing_low_smoothness}
-	\Ebb_{f_0}\bigl[1 - \wh{\varphi}\bigr] \leq \frac{C(1 + M^4)}{b^2},
+	\Ebb_{f_0}\bigl[1 - \wh{\varphi}\bigr] \leq \frac{C(1 + M^4)}{b^2}.
 	\end{equation} 
-	and \smash{$\Ebb_0[\wh{\varphi}] \leq 1/b^2$}.
-Note that the same results apply to \smash{$\wt{T}$}, since the thin-plate spline estimator \smash{$\wt{f}$} interpolates the responses $Y_1,\ldots,Y_n$ for $d>1$.
 	\item To compute the data-dependent threshold \smash{$\wh{t}_b$}, one must know all of the eigenvalues $\lambda_1,\ldots,\lambda_n$.

### 2021-03-09 manuscripts/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
Updated the AISTATS camera-ready version.

@@
 
-We adopt the standard setup in nonparametric regression, where we observe samples $(X_1,Y_1),\ldots,(X_n,Y_n)$ that are i.i.d.\ draws from the model 
+We adopt the standard nonparametric regression setup, where we observe samples $(X_1,Y_1),\ldots,(X_n,Y_n)$ that are i.i.d.\ draws from the model 
 \begin{equation}
 \label{eqn:signal_plus_noise_model}
-Y_i = f_0(X_i) + \varepsilon_i, \quad \epsilon_i \sim N(0,1), 
+Y_i = f_0(X_i) + \varepsilon_i, \quad \varepsilon_i \sim N(0,1), 
 \end{equation}
 where $\varepsilon_i$ is independent of $X_i$.
-Our goal is to perform statistical inference on the unknown regression function $f_0$, by which we mean either \emph{estimating} $f_0$ or more simply \emph{testing} whether $f_0 = 0$, i.e., whether there is any signal present.
+Our goal is to perform statistical inference on the unknown regression function $f_0$, by which we mean either \emph{estimating} $f_0$ or \emph{testing} whether $f_0 = 0$, i.e., whether there is any signal present.
 
 Laplacian smoothing \citep{smola2003} is a penalized least squares estimator, defined over a graph.
-Letting $G = (V,W)$ be a weighted undirected graph with vertices $V=\{1,\ldots,n\}$, associated with $\{X_1,\ldots,X_n\}$, the Laplacian smoothing estimator \smash{$\wh{f}$} is given by
+Letting $G = (V,W)$ be a weighted undirected graph with vertices $V=\{1,\ldots,n\}$, associated with $\{X_1,\ldots,X_n\}$, and $W \in \mathbb{R}^{n \times n}$ is the (weighted) adjacency matrix of the graph.
+the Laplacian smoothing estimator \smash{$\wh{f}$} is given by
 \begin{equation}
@@
 \section{SUMMARY OF RESULTS}
+\label{sec:summary}
 
@@
 \end{equation}
-where the optimization is performed over all functions for which the criterion is well-defined and finite (functions $f$ whose weak derivative $\nabla f$ in $L^2(\Xset)$).
-Analogously, for testing, we could use:
+where the optimization is performed over all continuous functions $f$ that have a weak derivative $\nabla f$ in $L^2(\Xset)$. Analogously, for testing, we could use:
 \begin{equation}
 \label{eqn:thin_plate_spline_test}
-\wt{T} = \| \wt{f} \|_n^2 = \frac{1}{n} \sum_{i=1}^n \wt{f}(X_i)^2.
+\wt{T} = \| \wt{f} \|_n^2 := \frac{1}{n} \sum_{i=1}^n \wt{f}(X_i)^2.
 \end{equation}
 The penalty term in \eqref{eqn:thin_plate_spline} leverages the assumption that $f_0$ has a smooth derivative in a seemingly natural way.
-Indeed, the estimator \smash{$\wt{f}$} and statistic \smash{$\wt{T}$} are well-known: for $d = 1$, \smash{$\wt{f}$} is the familiar \emph{smoothing spline}, and for $d>1$, it is type of a \emph{thin-plate spline}.
+Indeed, the estimator \smash{$\wt{f}$} and statistic \smash{$\wt{T}$} are well-known: for $d = 1$, \smash{$\wt{f}$} is the familiar \emph{smoothing spline}, and for $d>1$, it is a type of \emph{thin-plate spline}.
 The statistical properties of smoothing and thin-plate splines are well-understood \citep{vandergeer2000,liu2019}.
@@
 For sequences $a_n,b_n$, we write $a_n \lesssim b_n$ to mean $a_n \leq Cb_n$ for a constant $C>0$ and large enough $n$, and $a_n \asymp b_n$ to mean $a_n \lesssim b_n$ and $b_n \lesssim a_n$.
-Lastly, we use $a \wedge b = \max\{a,b\}$.
+Lastly, we use $a \wedge b = \min\{a,b\}$.
+%\sbcomment{This is wrong, should check where it appears.} \\
+%\agcomment{Resolved.}
 
@@
 \label{eqn:graph_Laplacian}
-\Lap_{n,r} = D - W.
+\Lap = D - W.
 \end{equation}
-We add subscripts $n,r$ to $\Lap$ and $G$ to emphasize that they depend on the design points $X_1,\ldots,X_n$ and the radius $r$.
-To avoid any confusion: henceforth when we speak of \smash{$\wh{f}$} and \smash{$\wh{T}$}, we will always assume that they are fit using the neighborhood graph Laplacian \smash{$L = L_{n,r}$}.
-Writing its eigendecomposition as \smash{$L = \sum_{k = 1}^{n} \lambda_k v_k v_k^\top$}, we always assume, by convention, ordered eigenvalues $0 = \lambda_1 \leq \cdots \leq \lambda_n$, and unit-norm eigenvectors.
-%we will make reference to these objects later on.
+We write the eigendecomposition of $L$ as \smash{$L = \sum_{k = 1}^{n} \lambda_k v_k v_k^\top$}, and we always assume, by convention, ordered eigenvalues $0 = \lambda_1 \leq \cdots \leq \lambda_n$, and unit-norm eigenvectors.
+%\sbcomment{Don't understand why you're switching between $L$ and $L_{n,r}$?} \\
+%\agcomment{Resolved.}
 
@@
 It's fine, I wouldn't change anything because it would involve explaining a lot more, I'm just pointing this out.
-(Throughout we assume $M \geq n^{-1/2}$, as otherwise the minimax rate is just the parametric rate $n^{-1}$.)
+(Throughout we assume $M \geq n^{-1/2}$, as otherwise the trivial estimator $\wh{f} = 0$ achieves smaller error than the parametric rate $n^{-1}$, and the problem doesn't fit well within the nonparametric setup.)
+%\sbcomment{I don't think the second part of this comment is correct the way you've defined the Sobolev class.} \\
+%\agcomment{Resolved, as I suggested in my Slack post.}
 
-As nonparametric hypothesis testing is (comparatively) less familiar than nonparametric estimation, we briefly summarize the main idea before stating the optimal error rate.
+As minimax rates in nonparametric hypothesis testing are (comparatively) less familiar than those in nonparametric estimation, we briefly summarize the main idea before stating the optimal error rate.
 In the goodness-of-fit testing problem, we ask for a test function---formally, a Borel measurable function $\phi$ taking values in $\{0,1\}$---which can distinguish between the hypotheses
@@
 In this paper, as in \citet{ingster2009}, we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\star} = 0$ and $\mc{F} = H^1(\Xset,M)$ is a first-order Sobolev ball.
+This is without loss of generality since our test statistic and its analysis are easily modified to handle the case when $f_0^{\star}$ is not $0$, by simply subtracting $f_0^{\star}(X_i)$ from each observation $Y_i$.
 
-The minimax critical radius is the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{F}_{\epsilon} := \mc{F} \cap \{f: \|f\|_{\Leb^2(\Xset)} \geq \epsilon\}$:
+%\sbcomment{Why isn't this without loss of generality?}\\
+%\agcomment{Don't know how to resolve.}
+
+%\sbcomment{Define a level-$\alpha$ test.
+You use it in~\eqref{eqn:type_I_error} for example.
+I also wouldn't re-use $\alpha$ for both Type-I and Type-II error.
+Later on you seem to use $b$ for Type-I error.
+Maybe use $b$ for Type-II and $\alpha$ for Type-I.}
+
+The Type I error of a test $\phi$ is $\mathbb{E}_0[\phi]$, and if $\mathbb{E}_0[\phi] \leq \alpha$ for a given $\alpha \in (0,1)$ we refer to $\phi$ as a level-$\alpha$ test.
+The worst-case risk of $\phi$ over $\mc{F}$ is
 \begin{equation*}
-\epsilon(\mc{F}) := \inf\biggl\{\epsilon > 0: \inf_{\phi} \sup_{f_0 \in \mc{F}_{\epsilon}} \Ebb_{f_0}[1 - \phi] \leq \alpha \biggr\},
+R_n(\phi, \mc{F}, \epsilon) := \sup\Bigl\{\mathbb{E}_{f_0}[1 - \phi]: f_0 \in \mc{F}, \|f_0\|_{\Leb^2(\mc{X})} > \epsilon\Bigr\},
 \end{equation*}
-where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Of course, the minimax critical radius $\epsilon$ depends on $\alpha$.
-However, we adopt the typical convention of treating $\alpha$ as a small but fixed positive constant; hence it will not affect the testing error rates, and we suppress it notationally.} See \citet{ingster82,ingster87,ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
+and for a given constant $b \geq 1$, the minimax critical radius $\epsilon(\mc{F})$ is the smallest value of $\epsilon$ such that some level-$\alpha$ test has worst-case risk of at most $1/b$.
+Formally,
+\begin{equation*}
+\epsilon(\mc{F}) := \inf\Bigl\{\epsilon > 0: \inf_{\phi} R_n(\phi,\mc{F},\epsilon) \leq 1/b \Bigr\},
+\end{equation*} 
+where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Clearly, the minimax critical radius $\epsilon$ depends on $\alpha$ and $b$.
+However, we adopt the typical convention of treating $\alpha$ and $b$ as small but fixed positive constants; hence it will not affect the testing error rates, and we suppress it notationally.} See \citet{ingster82,ingster87,ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
 
-Testing $f_0=0$ is an easier problem than estimating $f_0$, and hence the minimax testing critical radius over $H^1(\Xset,M)$ is smaller than the minimax estimation rate, for $1 \leq d < 4$ (see \citet{ingster2009}):
+Testing $f_0 = 0$ is an easier problem than estimating $f_0$, and hence the minimax testing critical radius over $H^1(\Xset,M)$ is smaller than the minimax estimation rate, for $1 \leq d < 4$ (see \citet{ingster2009}):
 \begin{equation}
@@
 	Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ has dimension $d < 4$ and $M \leq n^{1/d}$.
-Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume the neighborhood graph \smash{$G_{n,r}$} is computed with a kernel $K$ satifsying \ref{asmp:kernel}.
-There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_estimation}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$}, satisfies
+Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume the neighborhood graph \smash{$G_{n,r}$} is computed with a kernel $K$ satisfying \ref{asmp:kernel}.
+There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_estimation}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing} with \smash{$\rho = M^{-4/(2 + d)} (nr^{d + 2})^{-1} n^{-2/(2 + d)}$} satisfies
 	\begin{equation*}
@@
 	% AJG: Added.
-	\item The assumption $M \leq n^{1/d}$ ensures that the upper bound provided in the theorem is meaningful (i.e., ensures it is of most a constant order).
+	\item The assumption $M \leq n^{1/d}$ ensures that the upper bound provided in the theorem is meaningful (i.e., ensures it is of at most a constant order).
 	\item The lower bound on $r$ imposed in condition \ref{asmp:ls_kernel_radius_estimation} is compatible with practice, where by far the most common choice of radius is the connectivity threshold \smash{$r \asymp (\log(n)/n)^{1/d}$}, which makes \smash{$G_{n,r}$} as sparse as possible while still being connected, for maximum computational efficiency.
-The upper bound may seem a bit more mysterious---we need it for technical reasons to ensure that \smash{$\wh{f}$} does not overfit, but note that as a practical matter one rarely chooses $r$ to be so large anyway.
+The upper bound may seem a bit more mysterious---we need it for technical reasons to ensure that \smash{$\wh{f}$} does not overfit, but we note that as a practical matter one rarely chooses $r$ to be so large anyway.
 	\item It is possible to extend \smash{$\wh{f}$} to be defined on all of $\Xset$ and then evaluate the error of such an extension (as measured against $f_0$) in $\Leb^2(\Xset)$ norm.
 When \smash{$\wh{f}$} and $f_0$ are suitably smooth, tools from empirical process theory (see e.g., Chapter 14 of \citet{wainwright2019}) or approximation theory (e.g., Section 15.5 of \citet{johnstone2011}) guarantee that the $\Leb^2(\Xset)$ error is not too much greater than its in-sample counterpart.
-In fact, as shown in the supplement, if $f_0$ is Lipschitz smooth and we extend \smash{$\wh{f}$} to be piecewise constant over the Voronoi tesselation induced by $X_1,\ldots,X_n$, then the out-of-sample error \smash{$\|\wh{f}-f_0\|_{\Leb^2(\Xset)}$} is within a negligible factor of the in-sample error \smash{$\|\wh{f}-f_0\|_n$}.
+In fact, as we show in the supplement, if $f_0$ is Lipschitz smooth and we extend \smash{$\wh{f}$} to be piecewise constant over the Voronoi tessellation induced by $X_1,\ldots,X_n$, then the out-of-sample error \smash{$\|\wh{f}-f_0\|_{\Leb^2(\Xset)}$} is within a negligible factor of the in-sample error \smash{$\|\wh{f}-f_0\|_n$}.
 We leave analysis of the Sobolev case to future work.
-	\item When $f_0$ is Lipschitz smooth, we can also replace the factor of $\delta$ in the high probability bound by the sharper $\delta^2/n$.
+	\item When $f_0$ is Lipschitz smooth, we can also replace the factor of $\delta$ in the high probability bound by a factor of $\delta^2/n$, which is always smaller than $\delta$ when $\delta \in (0,1)$.
+	%\sbcomment{Too mysterious a comment. Why is this sharper?} \\
+	%\agcomment{Attempted to resolve.} 
 \end{itemize}
@@
 \end{theorem}
+%\sbcomment{I removed some negative sign above. Pls check.}\\
+%\agcomment{Good, thanks.}\\
 This mirrors the conclusions of \citet{sadhanala16} who investigate estimation rates of Laplacian smoothing over the $d$-dimensional grid graph.
@@
 
-For $b \geq 1$, define a threshold \smash{$\wh{t}_b$} as
+For a given $0 < \alpha < 1$, define a threshold \smash{$\wh{t}_{\alpha}$} as
 \begin{equation*}
-\wh{t}_{b} = \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}},
+\wh{t}_{\alpha} = \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2} + \frac{1}{n}\sqrt{\frac{2}{\alpha} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}},
 \end{equation*}
-where we recall $\lambda_k$ is the $k$th smallest eigenvalue of \smash{$\Lap_{n,r}$}. The Laplacian smoothing test is then simply
+where we recall $\lambda_k$ is the $k$th smallest eigenvalue of \smash{$L$}. The Laplacian smoothing test is then simply
 \begin{equation*}
-\wh{\varphi} = \1\bigl\{\wh{T} \leq \wh{t}_b\bigr\}.
+\wh{\varphi} = \1\bigl\{\wh{T} \leq \wh{t}_{\alpha}\bigr\}.
 \end{equation*} 
 %(Clearly \smash{$\wh{\varphi}$} depends on $b$, but we suppress this notationally.) 
-Here $b \geq 1$ is a tuning parameter selected by the user, with the choice driven by the tolerated Type I versus Type II error.
-In fact, the threshold \smash{$\wh{t}_b$} is precisely the right choice to control the Type I error of \smash{$\wh{\varphi}$}: we show in the supplement that
-\begin{equation}
-\label{eqn:type_I_error}
-\Ebb_0\bigl[\wh{\varphi}\bigr] \leq \frac{1}{b^2}.
-\end{equation}
-The next theorem shows that this test also has small Type II error, uniformly over all $f_0$ that are separated from $0$ by at least the critical radius given in \eqref{eqn:sobolev_space_testing_critical_radius}.
+We show in the supplement that $\wh{f}$ is a level-$\alpha$ test.
+In the next theorem, we upper bound the worst-case risk $R_n(\wh{\varphi},H^1(\mc{X},M),\epsilon)$ of $\wh{\varphi}$, whenever $\epsilon$ is at least (a constant times) the critical radius given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
 For this to hold, we will require a tighter range of scalings for the graph radius $r$.
@@
 \end{enumerate}
-We will also require that the radius of the graph Sobolev class not be too large. Precisely, we will require $M \leq M_{\max}(d)$, where we define
+We will also require that the radius of the Sobolev class not be too large. Precisely, we will require $M \leq M_{\max}(d)$, where we define
 \begin{equation*}
@@
 Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume \smash{$G_{n,r}$} is computed with a kernel $K$ satisfying \ref{asmp:kernel}.
-There are constants $N,C,c>0$ such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_testing}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$}, satisfies the following: for all $b \geq 1$, and all $f_0$ such that
+There are constants $N,C,C_1,c_1>0$ such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_testing}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$}, satisfies the following: for any $b \geq 1$, if
 	\begin{equation}
 	\label{eqn:laplacian_smoothing_testing}
-	\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2d/(4 + d)} n^{-4/(4 + d)},
+	\epsilon^2 \geq C M^{2d/(4 + d)} n^{-4/(4 + d)}\Bigl(b^2 + b\sqrt{\frac{1}{\alpha}}\Bigr) ,
 	\end{equation} 
-	the Type II error is upper bounded by
-	\begin{multline*}
-	\Ebb_{f_0}\bigl[1 - \wh{\varphi} \bigr] \leq \\
-	C\biggl(\frac{1}{b}\Bigl[1 + n^{-d/(4 + d)}M^{-2d/(4 + d)}\Bigr] + n\exp\bigl(-cnr^d\bigr)\biggr).
-	\end{multline*}
+	then the worst-case risk is upper bounded, $R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + C_1n\exp(-c_1nr^d)\bigr)$.
 \end{theorem}
+%\sbcomment{Personally, I wouldn't emphasize the Type II error so much in the Theorem (it is not tight and it's not an interesting quantity anyway), and write \eqref{eqn:laplacian_smoothing_testing} more directly as an upper bound on the critical radius (which is the interesting quantity).
+Otherwise you've introduced that terminology for no reason.}
 
@@
 On the other hand, if we explicitly assume that $f_0 \in \Leb^4(\Xset,M)$, then \citet{guerre02} show that the testing problem is characterized by a dimension-free lower bound $\epsilon^{2}(\Leb^4(\Xset,M)) \gtrsim n^{-1/2}$.
-Moreover, by training \smash{$\wh{f}$} to the interpolation limit, setting $\rho = 0$, the subsequent test \smash{$\wh{\varphi}$} will achieve (up to constants) this lower bound.
-That is, for any $f_0 \in \Leb^4(\Xset,M)$ such that \smash{$\|f_0\|_{\Leb^2(\Xset)}^2 \geq C b^2n^{-1/2}$}, we have both
+Moreover, by setting $\rho = 0$ so that the resulting estimator $\wh{f}$ interpolates the responses $Y_1,\ldots,Y_n$, the subsequent test \smash{$\wh{\varphi}$} will achieve (up to constants) this lower bound.
+That is, for any $f_0 \in \Leb^4(\Xset,M)$ such that \smash{$\|f_0\|_{\Leb^2(\Xset)}^2 \geq C(b^2 + \sqrt{1/\alpha})n^{-1/2}$}, we have that \smash{$\Ebb_0[\wh{\varphi}] \leq \alpha$} and
 	\begin{equation}
 	\label{eqn:laplacian_smoothing_testing_low_smoothness}
-	\Ebb_{f_0}\bigl[1 - \wh{\varphi}\bigr] \leq \frac{C(1 + M^4)}{b^2},
+	\Ebb_{f_0}\bigl[1 - \wh{\varphi}\bigr] \leq \frac{C(1 + M^4)}{b^2}.
 	\end{equation}
-	and \smash{$\Ebb_0[\wh{\varphi}] \leq 1/b^2$}.
-Note that the same results apply to \smash{$\wt{T}$}, since the thin-plate spline estimator \smash{$\wt{f}$} interpolates the responses $Y_1,\ldots,Y_n$ for $d>1$.
+	Note  
 	\item To compute the data-dependent threshold \smash{$\wh{t}_b$}, one must know all of the eigenvalues $\lambda_1,\ldots,\lambda_n$.
@@
 \end{equation*}
-and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies e.g., Dirichlet boundary conditions.} To be clear, this argument does not formally imply that $\wh{f}$ and $\wt{f}$ are close (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of functions with weak derivatives in $\Leb^2(\Xset)$).
-However, it does seem to suggest that the two estimators should behave somewhat similarly.
+and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies e.g., Dirichlet boundary conditions.} To be clear, this argument does not formally imply that the Laplacian eigenmaps estimator $\wh{f}$ and the thin-plate spline estimator $\wt{f}$ are close (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of continuous functions with weak derivatives in $\Leb^2(\Xset)$).
+But it does seem to suggest that the two estimators should behave somewhat similarly.
 
@@
 \begin{equation*}
-\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \underbrace{\vphantom{\sum_{k=1}^n}\frac{2\rho}{n} \bigl(f_0^\top \Lap_{n,r} f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{10}{n} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2}}_{\textrm{variance}},
+\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \underbrace{\vphantom{\sum_{k=1}^n}\frac{2\rho}{n} \bigl(f_0^\top L f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{10}{n} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2}}_{\textrm{variance}},
 \end{equation*}
-and likewise that $\wh{\varphi}$ has high power whenever
+and likewise that $\wh{\varphi}$ has small risk whenever
 \begin{equation*}
-\bigl\|f_0\bigr\|_n^2 \geq \underbrace{\vphantom{\sqrt{\sum_{k=1}^n}}\frac{2\rho}{n} \bigl(f_0^\top \Lap_{n,r} f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{4b}{n} \sqrt{\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}}}_{\textrm{variance}}.
+\bigl\|f_0\bigr\|_n^2 \geq \underbrace{\vphantom{\sqrt{\sum_{k=1}^n}}\frac{2\rho}{n} \bigl(f_0^\top L f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{4b}{n} \sqrt{\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}}}_{\textrm{variance}}.
 \end{equation*}
@@
 Lemma~\ref{lem:neighborhood_eigenvalue} leads to a sufficiently tight upper bound on the variance term whenever the radius $r$ is sufficiently small; precisely, when $r$ is upper bounded as in \ref{asmp:ls_kernel_radius_estimation} for estimation, or \ref{asmp:ls_kernel_radius_testing} for testing.
-The parameter $\rho$ is then tuned to minimize the sum of bias and variance, in the usual manner.
+The parameter $\rho$ is then chosen to minimize the sum of these upper bounds on bias and variance, as usual, and some straightforward calculations give Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing}.
+%\sbcomment{Should probably say something about the rest of the proof, i.e. is it straightforward to put these together to get the two Theorems?} \\
+%\agcomment{Attempted to resolve.}
 
@@
 
+% \sbcomment{I might point out that the Arias-Castro result is for density testing over Holder spaces, and that technically rates for regression testing on manifolds for Sobolev spaces are your contribution.}
+
 % RJT: Again, would venture a guess that these are stated results for Holder classes, not Sobolev classes.
@@
 
-On the other hand, a theory has been developed \citep{belkin03,belkin05,niyogi2013} establishing that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
-We contribute to this line of work by showing that under the manifold hypothesis, Laplacian smoothing achieves the sharper minimax rates over $H^1(\Xset)$.
+On the other hand, a theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+We contribute to this line of work by showing that under the manifold hypothesis, Laplacian smoothing achieves the tighter minimax rates over $H^1(\Xset)$.
 
@@
 	As in Theorem \ref{thm:laplacian_smoothing_testing}, but where $\Xset \subseteq \Rd$ is a manifold with intrinsic dimension $m < 4$, $M \leq M_{\max}(m)$, and the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}.
-There are constants $N,C,c>0$ such that for any $n \geq N$, and any $r$ as in \ref{asmp:ls_kernel_radius_testing_manifold}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$}, satisfies the following: for all $b \geq 1$, and $f_0$ such that
+	There are constants $N,C,C_1,c_1>0$ such that for any $n \geq N$, and any $r$ as in \ref{asmp:ls_kernel_radius_testing_manifold}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$}, satisfies the following: for any $b \geq 1$, if
 	\begin{equation}
 	\label{eqn:laplacian_smoothing_testing_manifold}
-	\bigl\|f_0\bigr\|_{\Leb^2(\Xset)}^2 \geq C b^2 M^{2m/(4 + m)} n^{-4/(4 + m)},
+	\epsilon^2 \geq C M^{2m/(4 + m)} n^{-4/(4 + m)}\Bigl(b^2 + b\sqrt{\frac{1}{\alpha}}\Bigr) ,
 	\end{equation} 
-	the Type II error is upper bounded by
-	\begin{multline*}
-	\Ebb_{f_0}\bigl[1 - \wh{\varphi} \bigr] \leq \\
-	C\biggl(\frac{1}{b}\Bigl[1 + n^{-m/(4 + m)}M^{-2m/(4 + m)}\Bigr] + n\exp\bigl(-cnr^m\bigr)\biggr).
-	\end{multline*}
+	then the worst-case risk is upper bounded, $R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + C_1n\exp(-cnr^m)\bigr)$.
 	% RJT: Alden, please verify that what I wrote is correct.
@@
 \end{theorem}
+% \sbcomment{Same comment as before on what to emphasize.}
 
-The proof of Theorems \ref{thm:laplacian_smoothing_estimation_manifold} and \ref{thm:laplacian_smoothing_testing_manifold} proceeds in a similar manner to that of Theorems \ref{thm:laplacian_smoothing_estimation1} and \ref{thm:laplacian_smoothing_testing}.
+The proofs of Theorems \ref{thm:laplacian_smoothing_estimation_manifold} and \ref{thm:laplacian_smoothing_testing_manifold} proceed in a similar manner to that of Theorems \ref{thm:laplacian_smoothing_estimation1} and \ref{thm:laplacian_smoothing_testing}.
 The key difference is that in the manifold setting, the equations \eqref{eqn:graph_sobolev_seminorm} and \eqref{eqn:neighborhood_eigenvalue} used to upper bound bias and variance will hold with $d$ replaced by $m$.
 
-We emphasize that little about $\Xset$ needs be known for Theorems~\ref{thm:laplacian_smoothing_estimation_manifold} and \ref{thm:laplacian_smoothing_testing_manifold} to hold.
+We emphasize that little about $\Xset$ need be known for Theorems~\ref{thm:laplacian_smoothing_estimation_manifold} and \ref{thm:laplacian_smoothing_testing_manifold} to hold.
 Indeed, all that is needed is the intrinsic dimension $m$, to properly tune $r$ and $\rho$ (from a theoretical point of view), and otherwise \smash{$\wh{f}$} and \smash{$\wh{\varphi}$} are computed without regard to $\Xset$.
-In contrast, the penalty in \eqref{eqn:thin_plate_spline} would have to be specially tailored to work in this setting, revealing another advantage of the discrete approach to the variational one.
+In contrast, the penalty in \eqref{eqn:thin_plate_spline} would have to be specially tailored to work in this setting, revealing another advantage of the discrete approach over the variational one.
 
@@
 In practice, it is more common to use a $k$-nearest-neighbor (kNN) graph than a neighborhood graph, due to the guaranteed connectivity and sparsity of the former; we suspect that by building on the work of \citet{calder2019}, one can show that our main results all hold under the kNN graph as well.
-In another direction, one can also generalize Laplacian smoothing by replacing the penalty \smash{$f^\top \Lap_{n,r} f$} with \smash{$f^\top \Lap_{n,r}^k f$}, for an integer $k > 1$.
-The hope is that this would then achieve minimax optimal rates over the higher-order Sobolev class $H^k(\Xset)$.
+In another direction, one can also generalize Laplacian smoothing by replacing the penalty \smash{$f^\top L f$} with \smash{$f^\top L^{s} f$}, for an integer $s > 1$.
+The hope is that this would then achieve minimax optimal rates over the higher-order Sobolev class $H^s(\Xset)$.
 % In the very special case of $k = 2$ and $d \leq 2$, we can show that this is indeed the case, but the general story--for all combinations of $k$ and $d$---remains beyond our reach.
+%\sbcomment{Not a huge fan of using $k$ in adjacent lines to refer to completely different things.}\\
+%\agcomment{Resolved.}
 
-\subsection*{Acknowledgements}
-
-AG and RJT were supported by ONR grant N00014-20-1-2787.
+\subsection*{Acknowledgments}
+AG and RJT were supported by ONR grant N00014-20-1-2787. AG and SB were supported by NSF grants DMS-1713003 and CCF-1763734.
 

### 2021-03-09 manuscripts/laplacian_smoothing/AISTATS/supplement_minimax_optimal_laplacian_smoothing.tex
Updated the AISTATS camera-ready version.

@@
 
-In this supplement, we provide proofs of all results found in the paper ``Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs''.
-Our main theorems (Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing_manifold}) all follow the same general proof strategy, centered on conditioning.
+\section{Preliminaries}
+\label{app:preliminaries}
+In this supplement, we provide complete proofs of all results found in the paper ``Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs''. 
+Our main theorems (Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing_manifold}) all follow the same general proof strategy of first establishing bounds in the fixed-design setup.
 In Section~\ref{sec:fixed_graph_error_bounds}, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to (functionals of) the graph $G$, and allow us to upper bound the error of $\wh{f}$ and $\wh{\varphi}$ conditional on the design $\{X_1,\ldots,X_n\} = \{x_1,\ldots,x_n\}$.
-In Sections~\ref{sec:graph_sobolev_seminorm}, \ref{sec:graph_eigenvalues}, \ref{sec:empirical_norm}, and \ref{sec:manifold} we give all our necessary probabilistic estimates on these functionals, for the particular random neighborhood graph $G = G_{n,r}$.
+In Sections~\ref{sec:graph_sobolev_seminorm}, \ref{sec:graph_eigenvalues}, \ref{sec:empirical_norm}, and \ref{sec:manifold} we develop all the necessary probabilistic estimates on these functionals, for the particular random neighborhood graph $G = G_{n,r}$.
 It is in these sections where we invoke our various assumptions on the distribution $P$ and regression function $f_0$.
@@
 In Lemma~\ref{lem:linear_smoother_fixed_graph_testing}, we upper bound the error of a test involving the statistic $\|\wc{f}\|_2^2 = {\bf Y}^{\top} S^2 {\bf Y}$.
+We will require that $S$ be a \emph{contraction}, meaning that it has operator norm no greater than $1$, $\|Sv\|_2 \leq \|v\|_2$ for all $v \in \Reals^n$.
 \begin{lemma}
 	\label{lem:linear_smoother_fixed_graph_testing}
-	Let $\wc{T} = Y^{\top} S^2 Y$ for a square, symmetric matrix $S \in \Reals^{n \times n}$. Define the threshold $\wc{t}_b$ to be 
-	\begin{equation}
-	\wc{t}_{b} := \norm{\lambdavec(S)}_2^2 + 2b \norm{\lambdavec(S)}_4^2
+	Let $\wc{T} = {\bf Y}^{\top} S^2 {\bf Y}$ for a square, symmetric matrix $S \in \Reals^{n \times n}$. Suppose $S$ is a contraction. Define the threshold $\wc{t}_{\alpha}$ to be 
+	\begin{equation}
+	\label{eqn:linear_smoother_fixed_graph_threshold}
+	\wc{t}_{\alpha} := \norm{\lambdavec(S)}_2^2 + \sqrt{\frac{2}{\alpha}} \norm{\lambdavec(S)}_4^2.
 	\end{equation}
@@
 		\label{eqn:linear_smoother_fixed_graph_testing_typeI}
-		\Pbb_0\bigl(\wc{T} > \wc{t}_b\bigr) \leq \frac{1}{b^2}
+		\Pbb_0\bigl(\wc{T} > \wc{t}_{\alpha}\bigr) \leq \alpha.
 		\end{equation}
@@
 \frac{1}{n}\bigl\|\wc{f} - f_0\bigr\|_2^2 & \leq \frac{2}{n}\Bigl(\bigl\|\wc{f} - \Ebb_{f_0}[\wc{f}]\bigr\|_2^2 + \bigl\|\Ebb_{f_0}[\wc{f}] - f_0\bigr\|_2^2\Bigr) \\ 
-& = \frac{2}{n}\Bigl(\bigl\|S\varepsilon\bigr\|_2^2 + \bigl\|(S - I)f_0\bigr\|_2^2\Bigr)
-\end{align*}
+& = \frac{2}{n}\Bigl(\bigl\|S\varepsilon\bigr\|_2^2 + \bigl\|(S - I)f_0\bigr\|_2^2\Bigr).
+\end{align*}
+%\sbcomment{General comment. Please punctuate math. I've done it till this point but please check everywhere.} \\
+%\agcomment{Resolved, I hope.} \\
 Writing $\norm{S\varepsilon}_2^2 = \sum_{k = 1}^{n} \lambda_k(S)^2 Z_k^2$, the claim follows from the result of \citet{laurent00} on concentration of $\chi^2$-random variables, which for completeness we restate in Lemma~\ref{lem:chi_square_bound}.
@@
 
-\textit{Bounding Type I and Type II error.} The upper bound \eqref{eqn:linear_smoother_fixed_graph_testing_typeI} follows immediately from \eqref{pf:linear_smoother_fixed_graph_testing_mean}, \eqref{pf:linear_smoother_fixed_graph_testing_var}, and Chebyshev's inequality.
-
-The bound \eqref{eqn:linear_smoother_fixed_graph_testing_typeII} also follows from Chebyshev's inequality, as can be seen by the following manipulations,
-\begin{equation*}
-\begin{aligned}
-\Pbb_{f_0}\Bigl(\wc{T} \leq \wc{t}_b\Bigr) & = \Pbb_{f_0}\Bigl(\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr] \leq \wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr) \\
-& \overset{(i)}{\leq} \Pbb_{f_0}\Bigl(\Bigl|\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr| \geq \Bigl|\wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr|\Bigr) \leq \frac{Var_{f_0}\bigl[\wc{T}\bigr]}{\bigl(\wc{t}_b - \Ebb_{f_0}\bigl[\wc{T}\bigr]\bigr)^2} \\
-& \overset{(ii)}{\leq} 4 \frac{\Var_{f_0}\bigl[\wc{T}\bigr]}{(f_0^{\top} S^2 f_0)^2} \\
-& \overset{(iii)}{\leq} \frac{32}{f_0^{\top} S^2 f_0} + \frac{4}{b^2} \\
-& \overset{(iv)}{\leq} \frac{8}{b} \Biggl(\sum_{k = 1}^{n}\bigl(\lambda_k(S)\bigr)^4\Biggr)^{-1/2} + \frac{4}{b^2}
-\end{aligned}
-\end{equation*}
-In the previous expression, $(i)$ and $(ii)$ follow since assumption \eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and equation \eqref{pf:linear_smoother_fixed_graph_testing_mean} together imply
-\begin{equation*}
-\wc{t}_b - \Ebb_{f_0}[\wc{T}] = f_0^{\top} S^2 f_0 - 2b\|{\bf \lambda}(S)\|_4^2 \geq \frac{1}{2}f_0^{\top} S^2 f_0
-\end{equation*}
-$(iii)$ follows from assumption \eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and the inequality \eqref{pf:linear_smoother_fixed_graph_testing_var}, and $(iv)$ follows from \eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}.
-
+\textit{Bounding Type I and Type II error.} The upper bound \eqref{eqn:linear_smoother_fixed_graph_testing_typeI} on Type I error follows immediately from \eqref{pf:linear_smoother_fixed_graph_testing_mean}, \eqref{pf:linear_smoother_fixed_graph_testing_var}, and Chebyshev's inequality.
+
+We now establish the upper bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeII} on Type II error.
+From assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}, we see that $f_0^{\top} S^2 f_0^{\top} - \wc{t}_{\alpha} \leq 0$.
+As a result,
+\begin{align*}
+\Pbb_{f_0}\Bigl(\wc{T} \leq \wc{t}_{\alpha}\Bigr) & = \Pbb_{f_0}\Bigl(\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr] \leq \wc{t}_{\alpha} - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr) \\ 
+& \leq \Pbb_{f_0}\Bigl(\Bigl|\wc{T} - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr| \geq \Bigl|\wc{t}_{\alpha} - \Ebb_{f_0}\bigl[\wc{T}\bigr]\Bigr|\Bigr) \\ 
+& \leq \frac{\Var_{f_0}\bigl[\wc{T}\bigr]}{\bigl(\wc{t}_{\alpha} - \Ebb_{f_0}\bigl[\wc{T}\bigr]\bigr)^2},
+\end{align*}
+where the last line follows from Chebyshev's inequality.
+Plugging in the expressions~\eqref{pf:linear_smoother_fixed_graph_testing_mean} and~\eqref{pf:linear_smoother_fixed_graph_testing_var} for the mean and variance of $\wc{T}$, as well as the definition of $\wc{t}_{\alpha}$ in~\eqref{eqn:linear_smoother_fixed_graph_threshold}, we obtain that
+\begin{equation}
+\label{pf:linear_smoother_fixed_graph_testing2}
+\Pbb_{f_0}\Bigl(\wc{T} \leq \wc{t}_{\alpha}\Bigr) \leq \frac{4\|\lambdavec(S)\|_4^4}{\bigl(f_0^{\top}S^2f_0 - \sqrt{2/\alpha} \norm{\lambdavec(S)}_4^2\bigr)^2} + \frac{8f_0^{\top}S^4f_0 }{\bigl(f_0^{\top}S^2f_0 - \sqrt{2/\alpha} \norm{\lambdavec(S)}_4^2\bigr)^2}.
+\end{equation}
+We now use the assumed lower bound $f_0^{\top} S^2 f_0 \geq (2\sqrt{2/\alpha} + 2b) \norm{\lambdavec(S)}_4^2$ to separately upper bound each of the two terms on the right hand side of~\eqref{pf:linear_smoother_fixed_graph_testing2}.
+It follows immediately that
+\begin{equation}
+\label{pf:linear_smoother_fixed_graph_testing2.5}
+\frac{4 \|\lambdavec(S)\|_4^4}{\bigl(f_0^{\top}S^2f_0 - \sqrt{2/\alpha} \norm{\lambdavec(S)}_4^2\bigr)^2} \leq \frac{1}{b^2},
+\end{equation}
+giving a sufficient upper bound on the first term. Now we upper bound the second term,
+\begin{equation}
+\label{pf:linear_smoother_fixed_graph_testing3}
+\frac{8f_0^{\top}S^4f_0 }{\bigl(f_0^{\top}S^2f_0 - \sqrt{2/\alpha} \norm{\lambdavec(S)}_4^2\bigr)^2} \leq \frac{32 f_0^{\top}S^4f_0 }{\bigl(f_0^{\top}S^2f_0\bigr)^2} \leq \frac{16}{b \|\lambdavec(S)\|_4^2} \frac{f_0^{\top}S^4f_0 }{f_0^{\top}S^2f_0} \leq \frac{16}{b \|\lambdavec(S)\|_4^2},
+\end{equation}
+where the final inequality is satisfied because $S$ is a contraction.
+Plugging~\eqref{pf:linear_smoother_fixed_graph_testing2.5} and~\eqref{pf:linear_smoother_fixed_graph_testing3} back into~\eqref{pf:linear_smoother_fixed_graph_testing2} then gives the desired result.\\
+%\sbcomment{I think the first equality should be the negative of what you wrote?} \\
+%\agcomment{You were correct.
+However I also rewrote the proof a bit to make it (hopefully) easier to follow.
+We also need that $S$ is a contraction---a condition that I had and then deleted at some point by accident--- and I added this to the statement of the Lemma.}
 \subsection{Analysis of Laplacian smoothing}
@@
 		\label{eqn:ls_fixed_graph_testing_typeI}
-		\Pbb_0\Bigl(\wh{T} > \wh{t}_b\Bigr) \leq \frac{1}{b^2}
+		\Pbb_0\Bigl(\wh{T} > \wh{t}_{\alpha}\Bigr) \leq \alpha.
 		\end{equation}
@@
 		\label{eqn:ls_fixed_graph_testing_critical_radius}
-		\frac{1}{n}\norm{f_0}_2^2 \geq \frac{2 \rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{4b}{n} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{1/2}
+		\frac{1}{n}\norm{f_0}_2^2 \geq \frac{2 \rho}{n} \bigl(f_0^{\top} \Lap f_0\bigr) + \frac{2\sqrt{2/\alpha} + 2b}{n} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{1/2},
 		\end{equation}
@@
 		\label{eqn:ls_fixed_graph_testing_typeII}
-		\Pbb_{f_0}\Bigl(\wh{T}(G) \leq \wh{t}_b\Bigr) \leq \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2}
+		\Pbb_{f_0}\Bigl(\wh{T}(G) \leq \wh{t}_{\alpha}\Bigr) \leq \frac{1}{b^2} + \frac{16}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \Biggr)^{-1/2}.
 		\end{equation}
@@
 \begin{aligned}
-\bigl\|(\wh{S} - I)f_0\bigr\|_2^2 & = f_0^{\top} \Lap^{1/2} \Lap^{-1/2}\bigl(\wh{S} - \Id\bigr) \Lap^{-1/2} \Lap^{1/2} f_0 \\
-& \leq f_0^{\top} \Lap f_0 \cdot \lambda_n\Bigl(\Lap^{-1/2}\bigl(\wh{S} - \Id\bigr)\Lap^{-1/2}\Bigr) \\
-& = f_0^{\top} \Lap f_0 \cdot \max_{k \in [n]} \biggl\{\frac{1}{\lambda_k} \Bigl(1 - \frac{1}{\rho\lambda_k + 1}\Bigr) \biggr\} \\
-& = f_0^{\top} \Lap f_0 \cdot \rho.
+\bigl\|(\wh{S} - I)f_0\bigr\|_2^2 & = f_0^{\top} \Lap^{1/2} \Lap^{-1/2}\bigl(\wh{S} - \Id\bigr)^2 \Lap^{-1/2} \Lap^{1/2} f_0 \\
+& \leq f_0^{\top} \Lap f_0 \cdot \lambda_n\Bigl(\Lap^{-1/2}\bigl(\wh{S} - \Id\bigr)^2 \Lap^{-1/2}\Bigr) \\
+& = f_0^{\top} \Lap f_0 \cdot \max_{k \in [n]} \biggl\{\frac{1}{\lambda_k} \Bigl(1 - \frac{1}{\rho\lambda_k + 1}\Bigr)^2 \biggr\} \\
+& \leq f_0^{\top} \Lap f_0 \cdot \rho.
 \end{aligned}
 \end{equation*} 
-In the above, we have written $\Lap^{-1/2}$ for the square root of the pseudoinverse of $\Lap$, and the maximum is over all indices $k$ such that $\lambda_k > 0$.
+In the above, we have written $\Lap^{-1/2}$ for the square root of the pseudoinverse of $\Lap$, the maximum is over all indices $k$ such that $\lambda_k > 0$, and the last inequality follows from the basic algebraic identity $1 - 1/(1 + x)^2 \leq 2 x$ for any $x > 0$.
 The claim of the Lemma then follows from Lemma~\ref{lem:linear_smoother_fixed_graph_estimation}.
@@
 & \leq f_0^{\top} \Lap f_0 \cdot  \lambda_{n}\biggl(\Lap^{-1/2}\Bigl(\Id - \wh{S}^2\Bigr) \Lap^{-1/2}\biggr) \\ 
-& \overset{(i)}{=}  f_0^{\top} \Lap f_0 \cdot \max_{k} \biggl\{ \frac{1}{\lambda_k} \Bigl(1 - \frac{1}{(\rho \lambda_k + 1)^2}\Bigr) \biggr\} \\
-& \overset{(ii)}{\leq} f_0^{\top} \Lap f_0 \cdot 2\rho,
+& = f_0^{\top} \Lap f_0 \cdot \max_{k} \biggl\{ \frac{1}{\lambda_k} \Bigl(1 - \frac{1}{(\rho \lambda_k + 1)^2}\Bigr) \biggr\} \\
+& \leq f_0^{\top} \Lap f_0 \cdot 2\rho,
 \end{aligned}
 \end{equation*}
----where in the above the maximum is over all indices $k$ such that $\lambda_k > 0$, and the last inequality follows from the basic algebraic identity $1 - 1/(1 + x)^2 \leq 2 x$ for any $x > 0$---we deduce that
-\begin{equation*}
-f_0^{\top} \wh{S}^2 f_0 \geq 2b \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{-1/2}.
+---where in the above the maximum is over all indices $k$ such that $\lambda_k > 0$---we deduce that
+\begin{equation*}
+f_0^{\top} \wh{S}^2 f_0 \geq  \Bigl(2\sqrt{\frac{2}{\alpha}} + 2b\Bigr) \cdot \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k + 1)^4} \biggr)^{-1/2}.
 \end{equation*} 
@@
 \label{pf:first_order_graph_sobolev_seminorm_1}
-\Ebb\Bigl[f^{\top} \Lap_{n,r} f\Bigr] = \frac{n(n - 1)}{2} \Ebb\biggl[\Bigl(f(X') - f(X)\Bigr)^2 K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr]
+\Ebb\Bigl[f^{\top} L f\Bigr] = \frac{n(n - 1)}{2} \Ebb\biggl[\Bigl(f(X') - f(X)\Bigr)^2 K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr],
 \end{equation}
@@
 	\begin{equation*}
-	f^{\top} \Lap_{n,r} f \leq \biggl(\frac{1}{\delta} + C_2\biggr) n^2 r^{d + 2} M^2
+	f^{\top} L f \leq \biggl(\frac{1}{\delta} + C_2\biggr) n^2 r^{d + 2} M^2.
 	\end{equation*}
@@
 \begin{equation*}
-\Pbb\biggl(\Bigl|f^{\top} \Lap_{n,r} f - \Ebb\bigl[f^{\top} \Lap_{n,r} f\bigr]\Bigr| \geq \frac{M^2}{\delta}n^2 r^{d + 2}\biggr) \leq C\frac{\delta^2}{n}
-\end{equation*}
-and since we have already upper bounded $\Ebb\bigl[f^{\top} \Lap_{n,r} f\bigr] \leq C_2 M^2 n^2 r^{d + 2}$, the proposition follows.
+\Pbb\biggl(\Bigl|f^{\top} L f - \Ebb\bigl[f^{\top} L f\bigr]\Bigr| \geq \frac{M^2}{\delta}n^2 r^{d + 2}\biggr) \leq C\frac{\delta^2}{n},
+\end{equation*}
+and since we have already upper bounded $\Ebb\bigl[f^{\top} L f\bigr] \leq C_2 M^2 n^2 r^{d + 2}$, the proposition follows.
 
@@
 In this section, we prove Lemma~\ref{lem:neighborhood_eigenvalue}, following the lead of~\citet{burago2014,trillos2019,calder2019}, who establish similar results with respect to a manifold without boundary.
-To prove this lemma, in  Theorem~\ref{thm:neighborhood_eigenvalue} we give estimates on the difference between eigenvalues of the graph Laplacian $L_{n,r}$ and eigenvalues of the weighted Laplace-Beltrami operator $\Delta_P$.
+To prove this lemma, in  Theorem~\ref{thm:neighborhood_eigenvalue} we give estimates on the difference between eigenvalues of the graph Laplacian $L$ and eigenvalues of the weighted Laplace-Beltrami operator $\Delta_P$.
 We recall $\Delta_P$ is defined as
 \begin{equation*}
-\Delta_Pf(x) := -\frac{1}{p(x)} \dive\bigl(p^2\nabla f\bigr)(x)
-\end{equation*}
-To avoid confusion, in this section we write $\lambda_k(G_{n,r})$ for the $k$th smallest eigenvalue of the graph Laplacian matrix $\Lap_{n,r}$ and $\lambda_k(\Delta_P)$ for the $k$th smallest eigenvalue of $\Delta_P$ \footnote{Under the assumptions~\ref{asmp:domain} and~\ref{asmp:density}, the operator $\Delta_P$ has a discrete spectrum; see \citet{garciatrillos18} for more details.}.
+\Delta_Pf(x) := -\frac{1}{p(x)} \dive\bigl(p^2\nabla f\bigr)(x).
+\end{equation*}
+To avoid confusion, in this section we write $\lambda_k(G_{n,r})$ for the $k$th smallest eigenvalue of the graph Laplacian matrix $L$ and $\lambda_k(\Delta_P)$ for the $k$th smallest eigenvalue of $\Delta_P$ \footnote{Under the assumptions~\ref{asmp:domain} and~\ref{asmp:density}, the operator $\Delta_P$ has a discrete spectrum; see \citet{garciatrillos18} for more details.}.
 Some other notation: throughout this section, we will write $A, A_0, A_1,\ldots$ and $a,a_0,a_1,\ldots$ for constants which may depend on $\Xset$, $d$, $K$, and $p$, but do not depend on $n$; we keep track of all such constants explicitly in our proofs.
 We let $L_K$ denote the Lipschitz constant of the kernel $K$.
-Finally, for notational ease we set $\theta$ and $\wt{\delta}$ to be the following (small) positive constants.
+Finally, for notational ease we set $\theta$ and $\wt{\delta}$ to be the following (small) positive numbers:
 \begin{equation}
@@
 \end{theorem}
-We note that Theorem~\ref{thm:neighborhood_eigenvalue} is not a strong enough result to imply that $\lambda_{k}(G_{n,r})$ is a consistent estimate of $\lambda_k(\Delta_P)$, i.e.
+Before moving forward to the proofs of Lemma~\ref{lem:neighborhood_eigenvalue} and Theorem~\ref{thm:neighborhood_eigenvalue}, it is worth being clear about the differences between Theorem~\ref{thm:neighborhood_eigenvalue} and the results of \citet{burago2014,trillos2019,calder2019}.
+First of all, the reason we cannot directly use the results of these works in the proof of Lemma~\ref{lem:neighborhood_eigenvalue} is that they all assume the domain $\mc{X}$ is without boundary, whereas for our results in Section~\ref{sec:minimax_optimal_laplacian_smoothing} we instead assume $\mc{X}$ has a (Lipschitz smooth) boundary.
+Fortunately, in this setting the high-level strategy shared by \citet{burago2014,trillos2019,calder2019} can still be used---indeed we follow it closely, as we summarize in Section~\ref{subsec:neighbrhood_eigenvalue_pf}.
+However, many calculations need to be redone, in order to account for points $x$ which are on or sufficiently close to the boundary of $\mc{X}$.
+For completeness and ease of reading, we provide a self-contained proof of Theorem~\ref{thm:neighborhood_eigenvalue}, but we comment where appropriate on connections between the technical results we use in this proof, and those derived in \citet{burago2014,trillos2019,calder2019}.
+
+On the other hand, we should also point out that unlike the results of~\citet{burago2014,trillos2019,calder2019}, Theorem~\ref{thm:neighborhood_eigenvalue} does not imply that $\lambda_{k}(G_{n,r})$ is a consistent estimate of $\lambda_k(\Delta_P)$, i.e.
 it does not imply that $|(nr^{d + 2})^{-1}\lambda_{k}(G_{n,r}) - \lambda_k(\Delta_P)| \to 0$ as $n \to \infty, r \to 0$.
-This is contrast to the results of~\citet{burago2014,trillos2019,calder2019}, whose results do imply consistency when the domain is assumed to be a manifold without boundary.
-While a more subtle analysis might imply consistency in our setting, the conclusion of Theorem~\ref{thm:neighborhood_eigenvalue}---that $\lambda_k(G_{n,r})/(nr^{d + 2}\lambda_k(\Delta_P))$ is bounded above and below by constants---suffices for our purposes.
-
-The bulk of this section is devoted to the proof of Theorem~\ref{thm:neighborhood_eigenvalue}.
-First, however, we show that our regularity under our conditions on $p$ and $\Xset$, Lemma~\ref{lem:neighborhood_eigenvalue} is a simple consequence of Theorem~\ref{thm:neighborhood_eigenvalue}.
+The key difficulty in proving consistency when $\mc{X}$ has a boundary can be summarized as follows: while at points $x \in \mc{X}$ satisfying $B(x,r) \subseteq \mc{X}$, the graph Laplacian $L$ is a reasonable approximation of the operator $\Delta_P$, at points $x$ near the boundary $L$ is known to approximate a different operator altogether \citep{belkin2012}.
+This is reminiscent of the boundary effects present in the analysis of kernel smoothing.
+We believe a more subtle analysis might imply convergence of eigenvalues in this setting.
+However, the conclusion of Theorem~\ref{thm:neighborhood_eigenvalue}---that $\lambda_k(G_{n,r})/(nr^{d + 2}\lambda_k(\Delta_P))$ is bounded above and below by constants that do not depend on $k$---suffices for our purposes.
+
+The bulk of the remainder of this section is devoted to the proof of Theorem~\ref{thm:neighborhood_eigenvalue}.
+First, however, we show that under our regularity conditions on $p$ and $\Xset$, Lemma~\ref{lem:neighborhood_eigenvalue} is a simple consequence of Theorem~\ref{thm:neighborhood_eigenvalue}.
 The link between the two is Weyl's Law.
@@
 
-\begin{proposition}[\textbf{c.f. Proposition 4.1 of \citet{calder2019}}]
+\begin{proposition}[\textbf{cf. Proposition 4.1 of \citet{calder2019}}]
 	\label{prop:dirichlet_energies}
-	With probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{m})$, we have the following.
+	With probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{d})$, we have the following.
 	\begin{enumerate}[(1)]
@@
 
-\begin{proposition}[\textbf{c.f. Proposition 4.2 of \citet{calder2019}}]
+\begin{proposition}[\textbf{cf. Proposition 4.2 of \citet{calder2019}}]
 	\label{prop:isometry}
-	With probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{m})$, we have the following.
+	With probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{d})$, we have the following.
 	\begin{enumerate}[(1)]
@@
 	\begin{equation*}
-	f^{\top} \Lap_{n,r} f \leq \frac{C}{\delta} n^2 r^{m + 2} |f|_{H^1(\Xset)}^2
+	f^{\top} L f \leq \frac{C}{\delta} n^2 r^{m + 2} |f|_{H^1(\Xset)}^2.
 	\end{equation*}
@@
 \paragraph{Proof of \eqref{eqn:laplacian_smoothing_testing_low_smoothness}.}
-When $\rho = 0$, the Laplacian smoother $\wh{f} = \mathbf{Y}$, the test statistic $\wh{T} = \frac{1}{n}\|\mathbf{Y}\|_2^2$, and the threshold $\wh{t}_b = 1 + 2bn^{-1/2}$.
+When $\rho = 0$, the Laplacian smoother $\wh{f} = \mathbf{Y}$, the test statistic $\wh{T} = \frac{1}{n}\|\mathbf{Y}\|_2^2$, and the threshold $\wh{t}_{\alpha} = 1 + n^{-1/2}\sqrt{2/\alpha}$.
 The expectation of $\wh{T}$ is
 \begin{equation*}
-\Ebb\bigl[\wh{T}\bigr] = \mathbb{E}\bigl[f_0^2(X)\bigr] + 1 \geq p_{\min} \norm{f_0}_{\Leb^2(\Xset)}^2 + 1
+\Ebb\bigl[\wh{T}\bigr] = \mathbb{E}\bigl[f_0^2(X)\bigr] + 1 \geq p_{\min} \norm{f_0}_{\Leb^2(\Xset)}^2 + 1.
 \end{equation*}
@@
 \begin{equation*}
-\norm{f_0}_{\Leb^2(X)}^2 \geq \frac{4b}{p_{\min}} n^{-1/2},
-\end{equation*}
-so that $E[\wh{T}] - \wh{t}_b \geq E[f_0^2(X)]/2$. Hence, by Chebyshev's inequality
-\begin{align*}
-\mathbb{P}_{f_0}\Bigl(\wh{T} \leq \wh{t}_b \Bigr) & \leq 4 \frac{\Var_{f_0}\bigl[\wh{T}\bigr]}{\mathbb{E}[f_0^2(X)]^2} \\
+\norm{f_0}_{\Leb^2(X)}^2 \geq \frac{2\sqrt{2/\alpha} + 2b}{p_{\min}} n^{-1/2},
+\end{equation*}
+so that $E[\wh{T}] - \wh{t}_{\alpha} \geq E[f_0^2(X)]/2$. Hence, by Chebyshev's inequality
+\begin{align*}
+\mathbb{P}_{f_0}\Bigl(\wh{T} \leq \wh{t}_{\alpha} \Bigr) & \leq 4 \frac{\Var_{f_0}\bigl[\wh{T}\bigr]}{\mathbb{E}[f_0^2(X)]^2} \\
 & \leq \frac{4}{n} \cdot \frac{ 3 + p_{\max}\bigl(M^4 + \|f_0\|_{\Leb^2(\Xset)}^2 \bigr)}{p_{\min}^2 \|f_0\|_{\Leb^2(\Xset)}^4} \\
-& \leq \frac{1}{16b^2}\Bigl(3 + \frac{4bp_{\max}}{p_{\min}n^{1/2}} + p_{\max}M^4\Bigr).
+& \leq \frac{1}{b^2}\Bigl(3 + \frac{4bp_{\max}}{p_{\min}n^{1/2}} + p_{\max}M^4\Bigr).
 \end{align*}
@@
 \begin{equation*}
-\Pbb\biggl(D_{n,r}(x) \geq (1 + \delta) nr^d p_{\max}\biggr) \leq 2\exp\biggl(-\frac{\delta^2nr^da_0p_{\min}}{3[K(0)]^2}\biggr),
-\end{equation*}
-and applying a union bound, we get that
-\begin{equation*}
-\Pbb\biggl(\max_{i = 1,\ldots,n}D_{n,r}(X_i) \geq (1 + \delta) nr^d p_{\max}\biggr) \leq 2n\exp\biggl(-\frac{\delta^2nr^da_0p_{\min}}{3[K(0)]^2}\biggr);
-\end{equation*}
-taking $\delta = 1$ gives the claimed upper bound.
+\Pbb\biggl(D_{n,r}(x) \geq (1 + \delta) nr^d p_{\max}\biggr) \leq 2\exp\biggl(-\frac{\delta^2nr^da_0p_{\min}}{3[K(0)]^2}\biggr).
+\end{equation*}
+Applying a union bound, we get that
+\begin{equation*}
+\Pbb\biggl(\max_{i = 1,\ldots,n}D_{n,r}(X_i) \geq (1 + \delta) nr^d p_{\max}\biggr) \leq 2n\exp\biggl(-\frac{\delta^2nr^da_0p_{\min}}{3[K(0)]^2}\biggr),
+\end{equation*}
+and taking $\delta = 1$ gives the claimed upper bound.
 

### 2021-03-09 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
Updated the AISTATS camera-ready version.

@@
 \label{eqn:signal_plus_noise_model}
-Y_i = f_0(X_i) + \varepsilon_i, \quad \epsilon_i \sim N(0,1), 
+Y_i = f_0(X_i) + \varepsilon_i, \quad \varepsilon_i \sim N(0,1), 
 \end{equation}
@@
 When \smash{$\wh{f}$} and $f_0$ are suitably smooth, tools from empirical process theory (see e.g., Chapter 14 of \citet{wainwright2019}) or approximation theory (e.g., Section 15.5 of \citet{johnstone2011}) guarantee that the $\Leb^2(\Xset)$ error is not too much greater than its in-sample counterpart.
-In fact, as shown in Appendix~\ref{sec:graph_sobolev_seminorm}, if $f_0$ is Lipschitz smooth and we extend \smash{$\wh{f}$} to be piecewise constant over the Voronoi tesselation induced by $X_1,\ldots,X_n$, then the out-of-sample error \smash{$\|\wh{f}-f_0\|_{\Leb^2(\Xset)}$} is within a negligible factor of the in-sample error \smash{$\|\wh{f}-f_0\|_n$}.
+In fact, as shown in Appendix~\ref{sec:graph_sobolev_seminorm}, if $f_0$ is Lipschitz smooth and we extend \smash{$\wh{f}$} to be piecewise constant over the Voronoi tessellation induced by $X_1,\ldots,X_n$, then the out-of-sample error \smash{$\|\wh{f}-f_0\|_{\Leb^2(\Xset)}$} is within a negligible factor of the in-sample error \smash{$\|\wh{f}-f_0\|_n$}.
 We leave analysis of the Sobolev case to future work.
-	\item When $f_0$ is Lipschitz smooth, we can also replace the factor of $\delta$ in the high probability bound by a factor of $\delta^2/n$, which is always smaller than $\delta$ when $\delta \in (0,1)$.\\
+	\item When $f_0$ is Lipschitz smooth, we can also replace the factor of $\delta$ in the high probability bound by a factor of $\delta^2/n$, which is always smaller than $\delta$ when $\delta \in (0,1)$.
 	%\sbcomment{Too mysterious a comment. Why is this sharper?} \\
@@
 
-We emphasize that little about $\Xset$ needs be known for Theorems~\ref{thm:laplacian_smoothing_estimation_manifold} and \ref{thm:laplacian_smoothing_testing_manifold} to hold.
+We emphasize that little about $\Xset$ need be known for Theorems~\ref{thm:laplacian_smoothing_estimation_manifold} and \ref{thm:laplacian_smoothing_testing_manifold} to hold.
 Indeed, all that is needed is the intrinsic dimension $m$, to properly tune $r$ and $\rho$ (from a theoretical point of view), and otherwise \smash{$\wh{f}$} and \smash{$\wh{\varphi}$} are computed without regard to $\Xset$.
@@
 
-\subsection*{Acknowledgements}
+\subsection*{Acknowledgments}
 AG and RJT were supported by ONR grant N00014-20-1-2787. AG and SB were supported by NSF grants DMS-1713003 and CCF-1763734.
 
-\bibliographystyle{unsrtnat}
+\bibliographystyle{plainnat}
 \bibliography{../../../graph_regression_bibliography}

### 2021-03-09 manuscripts/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
One last run through for typos.

@@
 Laplacian smoothing \citep{smola2003} is a penalized least squares estimator, defined over a graph.
-Letting $G = (V,W)$ be a weighted undirected graph with vertices $V=\{1,\ldots,n\}$, associated with $\{X_1,\ldots,X_n\}$, and $W \in \mathbb{R}^{n \times n}$ is the (weighted) adjacency matrix of the graph.
-the Laplacian smoothing estimator \smash{$\wh{f}$} is given by
+Let $G = (V,W)$ be a weighted undirected graph, where the vertices $V=\{1,\ldots,n\}$ are associated with $\{X_1,\ldots,X_n\}$, and $W \in \mathbb{R}^{n \times n}$ is the (weighted) adjacency matrix of the graph.
+The Laplacian smoothing estimator \smash{$\wh{f}$} is given by
 \begin{equation}
@@
 For these reasons, a body of work has emerged that analyzes the statistical properties of Laplacian smoothing, and graph-based methods more generally.
-Roughly speaking, this work can be divided into two categories, based on the perspective they adopt.
+Roughly speaking, these works can be divided into two categories, based on the perspective they adopt.
 \begin{itemize}
@@
 In this problem setting, tight upper bounds have been derived on the error of various graph-based methods (e.g., \citet{wang2016,hutter2016,sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}) and tests (e.g., \citet{sharpnack2010identifying,sharpnack2013b,sharpnack2013,sharpnack2015}), which certify that such procedures are \emph{optimal} over ``function'' classes (in quotes because these classes really model the $n$-dimensional vector of evaluations).
-The upside of this work is its generality: in this setting $G$ need not be a geometric graph, but in principle it could be any graph over $V=\{1,\ldots,n\}$.
+The upside of this work is its generality: in this setting $G$ need not be a geometric graph, but in principle could be any graph over $V=\{1,\ldots,n\}$.
 The downside is that, in the context of nonparametric regression, it is arguably not as natural to think of the evaluations of $f_0$ as exhibiting smoothness over some fixed pre-defined graph $G$, and more natural to speak of the smoothness of the function $f_0$ itself.
@@
 where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Clearly, the minimax critical radius $\epsilon$ depends on $\alpha$ and $b$.
-However, we adopt the typical convention of treating $\alpha$ and $b$ as small but fixed positive constants; hence it will not affect the testing error rates, and we suppress it notationally.} See \citet{ingster82,ingster87,ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
+However, we adopt the typical convention of treating $\alpha \in (0,1)$ and $b \geq 1$ as fixed positive constants; hence they will not affect the testing error rates, and we suppress them notationally.} See \citet{ingster82,ingster87,ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
 
@@
 	\end{equation}
-	Note  
-	\item To compute the data-dependent threshold \smash{$\wh{t}_b$}, one must know all of the eigenvalues $\lambda_1,\ldots,\lambda_n$.
+	\item To compute the data-dependent threshold \smash{$\wh{t}_{\alpha}$}, one must know all of the eigenvalues $\lambda_1,\ldots,\lambda_n$.
 Computing all these eigenvalues is far more expensive (cubic-time) than computing \smash{$\wh{T}$} in the first place (nearly-linear-time).
-But in practice we would not recommend using \smash{$\wh{t}_b$} anyway, and would instead we make the standard recommendation to calibrate via a permutation test \citep{hoeffding1952}.
+But in practice we would not recommend using \smash{$\wh{t}_{\alpha}$} anyway, and would instead we make the standard recommendation to calibrate via a permutation test \citep{hoeffding1952}.
 Recent work \cite{kim2020minimax}, has shown that in a variety of closely related settings, calibration of a test statistic via the permutation test often retains minimax-optimal power, and we expect similar results to hold for the Laplacian smoothing-based test statistic.
@@
 \end{equation*}
-and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies e.g., Dirichlet boundary conditions.} To be clear, this argument does not formally imply that the Laplacian eigenmaps estimator $\wh{f}$ and the thin-plate spline estimator $\wt{f}$ are close (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of continuous functions with weak derivatives in $\Leb^2(\Xset)$).
+and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies e.g., Dirichlet boundary conditions.} To be clear, this argument does not formally imply that the Laplacian smoothing estimator $\wh{f}$ and the thin-plate spline estimator $\wt{f}$ are close (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of continuous functions with weak derivatives in $\Leb^2(\Xset)$).
 But it does seem to suggest that the two estimators should behave somewhat similarly.
@@
 \begin{equation*}
-\bigl\|f_0\bigr\|_n^2 \geq \underbrace{\vphantom{\sqrt{\sum_{k=1}^n}}\frac{2\rho}{n} \bigl(f_0^\top L f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{4b}{n} \sqrt{\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}}}_{\textrm{variance}}.
+\bigl\|f_0\bigr\|_n^2 \geq \underbrace{\vphantom{\sqrt{\sum_{k=1}^n}}\frac{2\rho}{n} \bigl(f_0^\top L f_0\bigr)}_{\textrm{bias}} + \underbrace{\frac{2\sqrt{2/\alpha} + 2b}{n} \sqrt{\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}}}_{\textrm{variance}}.
 \end{equation*}
@@
 	As in Theorem \ref{thm:laplacian_smoothing_estimation1}, but where $\Xset \subseteq \Rd$ is a manifold with intrinsic dimension $m < 4$, the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}, and $M \leq n^{1/m}$.
-There are constants $N,C,C_1,c,c_1>0$ (not depending on $f_0$) such that for any \smash{$n \geq N$}, and any $r$ as in \ref{asmp:ls_kernel_radius_estimation_manifold}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + m)} (nr^{m + 2})^{-1} n^{-2/(2 + m)}$}, satisfies
+There are constants $N,C,c>0$ (not depending on $f_0$) such that for any \smash{$n \geq N$}, and any $r$ as in \ref{asmp:ls_kernel_radius_estimation_manifold}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + m)} (nr^{m + 2})^{-1} n^{-2/(2 + m)}$}, satisfies
 	\begin{equation*}
@@
 	\end{equation*}
-	with probability at least $1 - \delta -  C_1 n\exp(-cnr^m) - \exp(-c(M^2n)^{m/(2+m)})$.
+	with probability at least $1 - \delta -  Cn\exp(-cnr^m) - \exp(-c(M^2n)^{m/(2+m)})$.
 	% RJT: Alden, please verify that what I wrote is correct.
@@
 	As in Theorem \ref{thm:laplacian_smoothing_testing}, but where $\Xset \subseteq \Rd$ is a manifold with intrinsic dimension $m < 4$, $M \leq M_{\max}(m)$, and the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}.
-	There are constants $N,C,C_1,c_1>0$ such that for any $n \geq N$, and any $r$ as in \ref{asmp:ls_kernel_radius_testing_manifold}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$}, satisfies the following: for any $b \geq 1$, if
+	There are constants $N,C,c>0$ such that for any $n \geq N$, and any $r$ as in \ref{asmp:ls_kernel_radius_testing_manifold}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$}, satisfies the following: for any $b \geq 1$, if
 	\begin{equation}
@@
 	\end{equation} 
-	then the worst-case risk is upper bounded, $R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + C_1n\exp(-cnr^m)\bigr)$.
+	then the worst-case risk is upper bounded, $R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + Cn\exp(-cnr^m)$.
 	% RJT: Alden, please verify that what I wrote is correct.

### 2021-03-09 manuscripts/laplacian_smoothing/AISTATS/supplement_minimax_optimal_laplacian_smoothing.tex
One last run through for typos.

@@
 \textit{Proof of (1).}
-We begin by upper bounding $\bigl\|\wt{P}f\bigr\|_{n}$. By the Cauchy-Schwarz inequality and the bound on $\|\wt{p}_n - p\|_{\infty}$ in \eqref{eqn:optimal_transport_2},
-\begin{align*}
-\Bigl|\wt{P}f(X_i)\Bigr|^2 & = n^2 \Bigl|\int_{U_i} f(x) \wt{p}_n(x) \,dx\Bigr|^2 \\
+We begin by upper bounding $\bigl\|\wt{\mc{P}}f\bigr\|_{n}$. By the Cauchy-Schwarz inequality and the bound on $\|\wt{p}_n - p\|_{\infty}$ in \eqref{eqn:optimal_transport_2},
+\begin{align*}
+\Bigl|\wt{\mc{P}}f(X_i)\Bigr|^2 & = n^2 \Bigl|\int_{U_i} f(x) \wt{p}_n(x) \,dx\Bigr|^2 \\
 & \leq n \int_{U_i} \bigl|f(x)\bigr|^2 \wt{p}_n(x) \,dx \\
@@
 \label{pf:prop_isometry_1}
-\bigl\|\wt{P}f\bigr\|_{n}^2 \leq \biggl(1 + A_1(\theta + \wt{\delta})\biggr) \bigl\|f\bigr\|_{P}^2.
-\end{equation}
-Now, noticing that $\bigl\|\wt{P}f\bigr\|_{n} = \bigl\|\wt{P}^{\star}\wt{P}f\bigr\|_{\wt{P}_n}$, we can use the upper bound \eqref{pf:prop_isometry_1} to show that
+\bigl\|\wt{\mc{P}}f\bigr\|_{n}^2 \leq \biggl(1 + A_1(\theta + \wt{\delta})\biggr) \bigl\|f\bigr\|_{P}^2.
+\end{equation}
+Now, noticing that $\bigl\|\wt{\mc{P}}f\bigr\|_{n} = \bigl\|\wt{\mc{P}}^{\star}\wt{\mc{P}}f\bigr\|_{\wt{P}_n}$, we can use the upper bound \eqref{pf:prop_isometry_1} to show that
 \begin{align}
-\Bigl|\bigl\|\wt{P}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{P}^2\Bigr| & \leq \Bigl|\bigl\|\wt{P}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{\wt{P}_n}^2\Bigr| + \Bigl|\bigl\|f\bigr\|_{\wt{P}_n}^2 - \bigl\|f\bigr\|_{P}^2\Bigr| \nonumber \\
-& \overset{(i)}{\leq} \Bigl|\bigl\|\wt{P}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{\wt{P}_n}^2\Bigr|  + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2 \\
-& \overset{(ii)}{\leq} 2 \sqrt{1 + A_1(\theta + \wt{\delta})} \Bigl|\bigl\|\wt{P}f\bigr\|_{n} - \bigl\|f\bigr\|_{\wt{P}_n}\Bigr| \cdot \bigl\|f\bigr\|_{P} + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2 \nonumber \\
-& \leq 2 \sqrt{1 + A_1(\theta + \wt{\delta})} \bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\wt{P}_n} \cdot \bigl\|f\bigr\|_{P} + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2, \label{pf:prop_isometry_2}
+\Bigl|\bigl\|\wt{\mc{P}}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{P}^2\Bigr| & \leq \Bigl|\bigl\|\wt{\mc{P}}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{\wt{P}_n}^2\Bigr| + \Bigl|\bigl\|f\bigr\|_{\wt{P}_n}^2 - \bigl\|f\bigr\|_{P}^2\Bigr| \nonumber \\
+& \overset{(i)}{\leq} \Bigl|\bigl\|\wt{\mc{P}}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{\wt{P}_n}^2\Bigr|  + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2 \\
+& \overset{(ii)}{\leq} 2 \sqrt{1 + A_1(\theta + \wt{\delta})} \Bigl|\bigl\|\wt{\mc{P}}f\bigr\|_{n} - \bigl\|f\bigr\|_{\wt{P}_n}\Bigr| \cdot \bigl\|f\bigr\|_{P} + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2 \nonumber \\
+& \leq 2 \sqrt{1 + A_1(\theta + \wt{\delta})} \bigl\|\wt{\mc{P}}^{\star}\wt{\mc{P}}f - f\bigr\|_{\wt{P}_n} \cdot \bigl\|f\bigr\|_{P} + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2, \label{pf:prop_isometry_2}
 \end{align}
@@
 
-It remains to upper bound $\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\wt{P}_n}^2$. Noting that $\wt{P}^{\star}\wt{P}f$ is piecewise constant over the cells $U_i$, we have
-\begin{equation*}
-\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\wt{P}_n}^2 = \sum_{i = 1}^{n} \int_{U_i} \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx.
+It remains to upper bound $\bigl\|\wt{\mc{P}}^{\star}\wt{\mc{P}}f - f\bigr\|_{\wt{P}_n}^2$. Noting that $\wt{\mc{P}}^{\star}\wt{\mc{P}}f$ is piecewise constant over the cells $U_i$, we have
+\begin{equation*}
+\bigl\|\wt{\mc{P}}^{\star}\wt{\mc{P}}f - f\bigr\|_{\wt{P}_n}^2 = \sum_{i = 1}^{n} \int_{U_i} \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx.
 \end{equation*}
@@
 \begin{equation*}
-\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\wt{P}_n}^2 \leq A_3r^2 \wt{E}_r(f,\Xset) \leq  A_3 \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(\frac{C_5 p_{\max}^2}{p_{\min}^2}\Bigr) \cdot \sigma_{K}r^2 D_2(f),
+\bigl\|\wt{\mc{P}}^{\star}\wt{\mc{P}}f - f\bigr\|_{\wt{P}_n}^2 \leq A_3r^2 \wt{E}_r(f,\Xset) \leq  A_3 \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(\frac{C_5 p_{\max}^2}{p_{\min}^2}\Bigr) \cdot \sigma_{K}r^2 D_2(f),
 \end{equation*}

### 2021-03-09 manuscripts/laplacian_smoothing/arXiv/appendix.tex
One last run through for typos.

@@
 \label{pf:prop_isometry_1}
-\bigl\|\wt{P}f\bigr\|_{n}^2 \leq \biggl(1 + A_1(\theta + \wt{\delta})\biggr) \bigl\|f\bigr\|_{P}^2.
-\end{equation}
-Now, noticing that $\bigl\|\wt{P}f\bigr\|_{n} = \bigl\|\wt{P}^{\star}\wt{P}f\bigr\|_{\wt{P}_n}$, we can use the upper bound \eqref{pf:prop_isometry_1} to show that
+\bigl\|\wt{\mc{P}}f\bigr\|_{n}^2 \leq \biggl(1 + A_1(\theta + \wt{\delta})\biggr) \bigl\|f\bigr\|_{P}^2.
+\end{equation}
+Now, noticing that $\bigl\|\wt{\mc{P}}f\bigr\|_{n} = \bigl\|\wt{P}^{\star}\wt{P}f\bigr\|_{\wt{P}_n}$, we can use the upper bound \eqref{pf:prop_isometry_1} to show that
 \begin{align}
-\Bigl|\bigl\|\wt{P}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{P}^2\Bigr| & \leq \Bigl|\bigl\|\wt{P}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{\wt{P}_n}^2\Bigr| + \Bigl|\bigl\|f\bigr\|_{\wt{P}_n}^2 - \bigl\|f\bigr\|_{P}^2\Bigr| \nonumber \\
-& \overset{(i)}{\leq} \Bigl|\bigl\|\wt{P}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{\wt{P}_n}^2\Bigr|  + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2 \\
-& \overset{(ii)}{\leq} 2 \sqrt{1 + A_1(\theta + \wt{\delta})} \Bigl|\bigl\|\wt{P}f\bigr\|_{n} - \bigl\|f\bigr\|_{\wt{P}_n}\Bigr| \cdot \bigl\|f\bigr\|_{P} + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2 \nonumber \\
-& \leq 2 \sqrt{1 + A_1(\theta + \wt{\delta})} \bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\wt{P}_n} \cdot \bigl\|f\bigr\|_{P} + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2, \label{pf:prop_isometry_2}
+\Bigl|\bigl\|\wt{\mc{P}}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{P}^2\Bigr| & \leq \Bigl|\bigl\|\wt{\mc{P}}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{\wt{P}_n}^2\Bigr| + \Bigl|\bigl\|f\bigr\|_{\wt{P}_n}^2 - \bigl\|f\bigr\|_{P}^2\Bigr| \nonumber \\
+& \overset{(i)}{\leq} \Bigl|\bigl\|\wt{\mc{P}}f\bigr\|_{n}^2 - \bigl\|f\bigr\|_{\wt{P}_n}^2\Bigr|  + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2 \\
+& \overset{(ii)}{\leq} 2 \sqrt{1 + A_1(\theta + \wt{\delta})} \Bigl|\bigl\|\wt{\mc{P}}f\bigr\|_{n} - \bigl\|f\bigr\|_{\wt{P}_n}\Bigr| \cdot \bigl\|f\bigr\|_{P} + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2 \nonumber \\
+& \leq 2 \sqrt{1 + A_1(\theta + \wt{\delta})} \bigl\|\wt{\mc{P}}^{\star}\wt{\mc{P}}f - f\bigr\|_{\wt{P}_n} \cdot \bigl\|f\bigr\|_{P} + A_1(\theta + \wt{\delta}) \bigl\|f\bigr\|_{P}^2, \label{pf:prop_isometry_2}
 \end{align}

### 2021-03-09 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
One last run through for typos.

@@
 where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Clearly, the minimax critical radius $\epsilon$ depends on $\alpha$ and $b$.
-However, we adopt the typical convention of treating $\alpha$ and $b$ as small but fixed positive constants; hence it will not affect the testing error rates, and we suppress it notationally.} See \citet{ingster82,ingster87,ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
+However, we adopt the typical convention of treating $\alpha \in (0,1)$ and $b \geq 1$ as small but fixed positive constants; hence they will not affect the testing error rates, and we suppress them notationally.} See \citet{ingster82,ingster87,ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
 
@@
 	\end{equation} 
-	\item To compute the data-dependent threshold \smash{$\wh{t}_b$}, one must know all of the eigenvalues $\lambda_1,\ldots,\lambda_n$.
+	\item To compute the data-dependent threshold \smash{$\wh{t}_{\alpha}$}, one must know all of the eigenvalues $\lambda_1,\ldots,\lambda_n$.
 Computing all these eigenvalues is far more expensive (cubic-time) than computing \smash{$\wh{T}$} in the first place (nearly-linear-time).
-But in practice we would not recommend using \smash{$\wh{t}_b$} anyway, and would instead we make the standard recommendation to calibrate via a permutation test \citep{hoeffding1952}.
+But in practice we would not recommend using \smash{$\wh{t}_{\alpha}$} anyway, and would instead we make the standard recommendation to calibrate via a permutation test \citep{hoeffding1952}.
 Recent work \cite{kim2020minimax}, has shown that in a variety of closely related settings, calibration of a test statistic via the permutation test often retains minimax-optimal power, and we expect similar results to hold for the Laplacian smoothing-based test statistic.
@@
 	\end{equation*}
-	with probability at least $1 - \delta -  C_1 n\exp(-cnr^m) - \exp(-c(M^2n)^{m/(2+m)})$.
+	with probability at least $1 - \delta -  C n\exp(-cnr^m) - \exp(-c(M^2n)^{m/(2+m)})$.
 	% RJT: Alden, please verify that what I wrote is correct.
@@
 	\end{equation} 
-	then the worst-case risk is upper bounded, $R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + C_1n\exp(-c_1nr^m)\bigr)$.
+	then the worst-case risk is upper bounded, $R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + Cn\exp(-cnr^m)$.
 	% RJT: Alden, please verify that what I wrote is correct.

### 2021-03-10 manuscripts/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
Found one last typo.
Few trivial edits

@@
 Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius of the Sobolev ball is also optimal.
-Rates for thin-plate splines with $d \geq 2$ assume the estimator $\wt{f}$ interpolates the responses, $\wt{f}(X_i) = Y_i$ for $i = 1,\ldots,n$; see the discussion in Section~\ref{sec:summary}.
-We use ``$1$'' to indicate inconsistency (error not converging to 0).
-Furthermore, when $\Xset$ is an $m$-dimensional manifold embedded in $\Rd$, all Laplacian smoothing results hold with $d$ replaced by $m$, without any change to the method itself.}
+The rates for thin-plate splines with $d \geq 2$ assume the estimator \smash{$\wt{f}$} interpolates the responses, \smash{$\wt{f}(X_i) = Y_i$} for $i = 1,\ldots,n$; see the discussion in Section~\ref{sec:summary}.
+Here, we use ``$1$'' to denote inconsistency (error not converging to 0).
+Lastly, when $\Xset$ is an $m$-dimensional manifold embedded in $\Rd$, all Laplacian smoothing results hold with $d$ replaced by $m$, without any change to the method itself.}
 	\label{tbl:estimation_rates}
@@
 	\caption{Summary of testing rates over first-order Sobolev balls; black, red, and bold fonts are used as in Table~\ref{tbl:estimation_rates}.
-Rates for thin-plate splines with $d \geq 2$ assume the test statistic $\wt{T}$ is computed with respect to an estimator $\wt{f}$ that interpolates the responses, $\wt{f}(X_i) = Y_i$ for $i = 1,\ldots,n$.
+The rates for thin-plate splines with $d \geq 2$ assume the test statistic \smash{$\wt{T}$} is computed using an \smash{$\wt{f}$} that interpolates the responses, \smash{$\wt{f}(X_i) = Y_i$} for $i = 1,\ldots,n$.
 Rates for $d \geq 4$ assume that $f_0 \in \Leb^4(\Xset,M)$.
-When $\Xset$ is an $m$-dimensional manifold embedded in $\Rd$, all rates hold with $d$ replaced by $m$.}
+Lastly, when $\Xset$ is an $m$-dimensional manifold embedded in $\Rd$, all rates hold with $d$ replaced by $m$.}
 	\label{tbl:testing_rates}
@@
 \end{equation}
-We write the eigendecomposition of $L$ as \smash{$L = \sum_{k = 1}^{n} \lambda_k v_k v_k^\top$}, and we always assume, by convention, ordered eigenvalues $0 = \lambda_1 \leq \cdots \leq \lambda_n$, and unit-norm eigenvectors.
+We use \smash{$L = \sum_{k = 1}^{n} \lambda_k v_k v_k^\top$} for an eigendecomposition of $L$, and we always assume, by convention, ordered eigenvalues $0 = \lambda_1 \leq \cdots \leq \lambda_n$, and unit-norm eigenvectors.
 %\sbcomment{Don't understand why you're switching between $L$ and $L_{n,r}$?} \\
@@
 It's fine, I wouldn't change anything because it would involve explaining a lot more, I'm just pointing this out.
-(Throughout we assume $M \geq n^{-1/2}$, as otherwise the trivial estimator $\wh{f} = 0$ achieves smaller error than the parametric rate $n^{-1}$, and the problem doesn't fit well within the nonparametric setup.)
+(Throughout we assume $M \geq n^{-1/2}$, as otherwise the trivial estimator \smash{$\wh{f} = 0$} achieves smaller error than the parametric rate $n^{-1}$, and the problem does not fit well within the nonparametric setup.)
 %\sbcomment{I don't think the second part of this comment is correct the way you've defined the Sobolev class.} \\
@@
 \begin{equation*}
-\wh{\varphi} = \1\bigl\{\wh{T} \leq \wh{t}_{\alpha}\bigr\}.
+\wh{\varphi} = \1\bigl\{\wh{T} > \wh{t}_{\alpha}\bigr\}.
 \end{equation*} 
 %(Clearly \smash{$\wh{\varphi}$} depends on $b$, but we suppress this notationally.) 
-We show in the supplement that $\wh{f}$ is a level-$\alpha$ test.
-In the next theorem, we upper bound the worst-case risk $R_n(\wh{\varphi},H^1(\mc{X},M),\epsilon)$ of $\wh{\varphi}$, whenever $\epsilon$ is at least (a constant times) the critical radius given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
+We show in the supplement that \smash{$\wh{f}$} is a level-$\alpha$ test.
+In the next theorem, we upper bound the worst-case risk \smash{$R_n(\wh{\varphi},H^1(\mc{X},M),\epsilon)$} of \smash{$\wh{\varphi}$}, whenever $\epsilon$ is at least (a constant times) the critical radius given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
 For this to hold, we will require a tighter range of scalings for the graph radius $r$.
@@
 	\label{thm:laplacian_smoothing_testing}
-	Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ with $d < 4$, and where $M \leq M_{\max}(d)$.
+	Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ with $d < 4$, and $M \leq M_{\max}(d)$.
 Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume \smash{$G_{n,r}$} is computed with a kernel $K$ satisfying \ref{asmp:kernel}.
-There are constants $N,C,C_1,c_1>0$ such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_testing}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$}, satisfies the following: for any $b \geq 1$, if
+There exist constants $N,C,C_1,c_1>0$ such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_testing}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$}, satisfies the following: for any $b \geq 1$, if
 	\begin{equation}
 	\label{eqn:laplacian_smoothing_testing}
-	\epsilon^2 \geq C M^{2d/(4 + d)} n^{-4/(4 + d)}\Bigl(b^2 + b\sqrt{\frac{1}{\alpha}}\Bigr) ,
+	\epsilon^2 \geq C M^{2d/(4 + d)} n^{-4/(4 + d)}\biggl(b^2 + b\sqrt{\frac{1}{\alpha}}\biggr) ,
 	\end{equation} 
-	then the worst-case risk is upper bounded, $R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + C_1n\exp(-c_1nr^d)\bigr)$.
+	then the worst-case risk satisfies the upper bound: \smash{$R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + C_1n\exp(-c_1nr^d)$}.
 \end{theorem}
@@
 On the other hand, if we explicitly assume that $f_0 \in \Leb^4(\Xset,M)$, then \citet{guerre02} show that the testing problem is characterized by a dimension-free lower bound $\epsilon^{2}(\Leb^4(\Xset,M)) \gtrsim n^{-1/2}$.
-Moreover, by setting $\rho = 0$ so that the resulting estimator $\wh{f}$ interpolates the responses $Y_1,\ldots,Y_n$, the subsequent test \smash{$\wh{\varphi}$} will achieve (up to constants) this lower bound.
+Moreover, by setting $\rho = 0$ so that the resulting estimator \smash{$\wh{f}$} interpolates the responses $Y_1,\ldots,Y_n$, the subsequent test \smash{$\wh{\varphi}$} will achieve (up to constants) this lower bound.
 That is, for any $f_0 \in \Leb^4(\Xset,M)$ such that \smash{$\|f_0\|_{\Leb^2(\Xset)}^2 \geq C(b^2 + \sqrt{1/\alpha})n^{-1/2}$}, we have that \smash{$\Ebb_0[\wh{\varphi}] \leq \alpha$} and
@@
 \end{equation*}
-and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies e.g., Dirichlet boundary conditions.} To be clear, this argument does not formally imply that the Laplacian smoothing estimator $\wh{f}$ and the thin-plate spline estimator $\wt{f}$ are close (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of continuous functions with weak derivatives in $\Leb^2(\Xset)$).
+and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies e.g., Dirichlet boundary conditions.} To be clear, this argument does not formally imply that the Laplacian smoothing estimator \smash{$\wh{f}$} and the thin-plate spline estimator \smash{$\wt{f}$} are close (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of continuous functions with weak derivatives in $\Leb^2(\Xset)$).
 But it does seem to suggest that the two estimators should behave somewhat similarly.
@@
 \end{equation*}
-and likewise that $\wh{\varphi}$ has small risk whenever
+and likewise that \smash{$\wh{\varphi}$} has small risk whenever
 \begin{equation*}
@@
 
-On the other hand, a theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+On the other hand, a theory has been developed \citep{belkin03,belkin05,niyogi2008finding,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
 We contribute to this line of work by showing that under the manifold hypothesis, Laplacian smoothing achieves the tighter minimax rates over $H^1(\Xset)$.
@@
 \end{theorem}
-In a similar vein, we obtain results for manifold adaptive testing under the following condition on the graph radius parameter:
+In a similar vein, we obtain results for manifold adaptive testing under the following condition on the graph radius parameter.
 \begin{enumerate}[label=(R\arabic*)]
@@
 	\label{eqn:laplacian_smoothing_testing_manifold}
-	\epsilon^2 \geq C M^{2m/(4 + m)} n^{-4/(4 + m)}\Bigl(b^2 + b\sqrt{\frac{1}{\alpha}}\Bigr) ,
+	\epsilon^2 \geq C M^{2m/(4 + m)} n^{-4/(4 + m)}\biggl(b^2 + b\sqrt{\frac{1}{\alpha}}\biggr) ,
 	\end{equation} 
-	then the worst-case risk is upper bounded, $R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + Cn\exp(-cnr^m)$.
+	then the worst-case risk satisfies the upper bound: \smash{$R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + Cn\exp(-cnr^m)$}.
 	% RJT: Alden, please verify that what I wrote is correct.
@@
 
-We emphasize that little about $\Xset$ need be known for Theorems~\ref{thm:laplacian_smoothing_estimation_manifold} and \ref{thm:laplacian_smoothing_testing_manifold} to hold.
+We emphasize that little about $\Xset$ needs be known for Theorems~\ref{thm:laplacian_smoothing_estimation_manifold} and \ref{thm:laplacian_smoothing_testing_manifold} to hold.
 Indeed, all that is needed is the intrinsic dimension $m$, to properly tune $r$ and $\rho$ (from a theoretical point of view), and otherwise \smash{$\wh{f}$} and \smash{$\wh{\varphi}$} are computed without regard to $\Xset$.

### 2021-03-10 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Completed out-of-sample extension section.

@@
 \end{enumerate}
-Under assumptions~\ref{asmp:domain}-\ref{asmp:sobolev_radius}, the minimax estimation rate over $\mc{H} = H^s(\mc{X};M)$ or $\mc{H} = H_0^s(\mc{X};M)$) is (see e.g. \citet{tsybakov2008_book})
+Under assumptions~\ref{asmp:domain}-\ref{asmp:sobolev_radius}, the minimax estimation rate over $\mc{H} = H^s(\mc{X};M)$ or $\mc{H} = H_0^s(\mc{X};M)$) is (see e.g. \citet{tsybakov08})
 \begin{equation}
@@
 \begin{itemize}
-	\item Is there a better citation than \citet{tsybakov2008_book} for the minimax rate of estimation~\eqref{eqn:minimax_estimation_rate}?
-Tsybakov considers the univariate case with fixed, equispaced design.
+	\item Is there a better citation than \citet{tsybakov08} for the minimax rate of estimation~\eqref{eqn:minimax_estimation_rate}? Tsybakov considers the univariate case with fixed, equispaced design.
 	\item I copy pasted the testing part directly from our AISTATS paper, then made a few necessary changes. Is that bad?
@@
 In the ``sub-critical'' regime $2s < d$, the Sobolev space $H^s(\mc{X})$ is quite irregular.
-For instance, it is not a Reproducing Kernel Hilbert Space, nor does it embed into any H\"{o}lder space.
+For instance, it is not a Reproducing Kernel Hilbert Space (RKHS), nor does it embed into any H\"{o}lder space.
 As a result, certain versions of the nonparametric regression problem are ill-posed---for instance when loss is measured in $\Leb^{\infty}$ norm, or when the design points $\{X_1,\ldots,X_n\}$ are assumed to be fixed.
@@
 When $4 < d \leq 4s$, we get the rate~\textcolor{red}{(...)} Recall that the rate $n^{-4s/(4s + d)}$ is optimal whenever $d \leq 4s$.
-Thus when $4s > d$, our upper bounds do not imply that $\varphi$ is an optimal test.
+Thus when $4 < d < 4s$, our upper bounds do not imply that $\varphi$ is an optimal test.
 We do not know whether this is due to looseness in our proof techniques, or a problem with the test itself, and leave this for future work.
@@
 \begin{itemize}
-	\item \textbf{Tuning parameters by sample-splitting.} 
-	\item \textbf{Other extensions.} We consider kernel smoothing because it is a simple and intuitive procedure that does not rely on knowledge of the domain $\mc{X}$ or distribution $P$.
+	\item \textbf{Tuning parameters by sample-splitting.} By extending the estimator $\wh{f}$ to be defined out-of-sample, extension of Laplacian Eigenmaps by kernel smoothing gives us the option of selecting the parameters $\varepsilon$ and $K$ in a data-dependent manner---e.g.
+by sample splitting or cross-validation.
+This is by far more the most common practical recommendation for selecting hyper-parameters \textcolor{red}{(references)}, and addresses a serious limitation of Laplacian Eigenmaps.
+Moreover, it is known \textcolor{red}{(references)} that choosing hyper-parameters by sample splitting results in estimators that optimality adapt to the order of regularity $s$---that is, that are rate-optimal (up to $\log$ factors), even when $s$ is unknown.
+We believe that similar arguments will imply that $T_{h,n}\wh{f}$ is adaptive in this sense, though we do not pursue the details.
+	\item \textbf{Other extensions.} We consider extension by kernel smoothing because it is a simple and statistically optimal procedure that does not require any knowledge of the domain $\mc{X}$ or distribution $P$.
 There exist (many) other approaches to extending (or interpolating) a function $f$ based on the evaluations $\{f(X_1),\ldots,f(X_n)\}$, and we now comment on a few of these.
@@
 	\end{equation*}
-	A particularly popular version of this general approach is when $\mc{F}$ is an RKHS, which encompasses thin-plate spline interpolation $\mc{F} = H^s(\mc{X})$ for $s > d/2$ as a special case.
-This approach works well when $f$ is close to a function $u \in \mc{F}$ with reasonably small norm, which holds true by assumption  but as already discussed when $2s \leq d$ the Sobolev space $H^s(P)$ is not a Reproducing Kernel Hilbert Space.
-Another
-	\item \textbf{Comparison with kernel smoothing.} 
+	A particularly popular version of this general approach takes $\mc{F}$ to be an RKHS, which encompasses thin-plate spline interpolation (where $\mc{F} = H^s(P)$ for $s > d/2$) as a special case.
+Naturally, this approach works well when $f$ is close to a function $u \in \mc{F}$ with reasonably small norm.
+This holds true when $f \in H^s(P;M)$ and $s > d/2$, but as already discussed when $s \leq d/2$ the Sobolev space $H^s(P)$ is not an RKHS, and in fact when $s \leq d/2$ thin-plate spline interpolation is an ill-posed problem \citep{green93}.
+Another, arguably simpler approach is to extend $\wh{f}$ to be piecewise constant on the Voronoi tessellation induced by $X_1,\ldots,X_n$, or equivalently to perform $1$-nearest neighbors regression on $\wh{f}$.
+However, this approach is theoretically optimal only when the regression function $f_0 \in C^1(P)$, in contrast to the kernel smoothing method we propose and study.
+	%Finally, the Nystr\"{o}m method extends  eigenvectors $v_k$ to functions $f_k: \mc{X} \to \Reals$ as follows,
+	%\begin{equation*}
+	%L_{n,\varepsilon}f_k = \lambda_k f_k~~\textrm{on $\mc{X}$, such that}~~f_k(X_i) = %v_k(X_i)~~\textrm{for $i = 1,\ldots,n$.}
+	%\end{equation*}
+	%Here $L_{n,\varepsilon}$ should be thought of as an operator acting on functions $f \in C(\mc{X})$ as follows,
+	%\begin{equation*}
+	%L_{n,\varepsilon}f(x) = \sum_{i = 1}^{n} (f(x) - f(X_i))\eta\Bigl(\frac{\|X_i - x\|}{\varepsilon}\Bigr)
+	%\end{equation*}
+	% The Nystr\"{o}m approach to extension makes intriguing (re-)use of the graph Laplacian $L_{n,\varepsilon}$ and its eigenvectors $v_k$.
+Similar methods have been analyzed in the context of kernel ridge regression (references), but little is known about its application to graph Laplacians.
 \end{itemize}
-
-\paragraph{Why not just kernel smoothing?}
-On the other hand, the allusion to kernel smoothing raises a very natural question: why should we prefer to first compute $\wh{f}$ and then smooth it out, as opposed to simply performing kernel smoothing over the original responses $Y$, that?
-
-We now answer this question, by \textcolor{red}{(...)}.
 
@@
 \label{sec:discussion}
+
+\paragraph{Comparison with kernel smoothing.}
+It is natural to ask whether the two stage estimator $T_{h,n}\wh{f}$ defined in Section~\ref{sec:out_of_sample} has any advantage over the simpler approach of directly kernel smoothing the responses, i.e.
+using the estimator $T_{h,n}Y$ (possibly for a different choice of $h$.) In Appendix~\textcolor{red}{(?)}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\|f_0 - T_{h,n}\wh{f}\|_P^2 \ll \inf_{h'} \|f_0 - T_{h',n}Y\|_P^2$.
+This is possible because Laplacian eigenmaps induces a completely different bias than kernel smoothing, which depends on the relationship between $f_0$ and $p$.
+When $f_0$ and $p$ satisfy the so-called \emph{cluster} assumption--- i.e.
+$f_0$ is piecewise constant in high-density regions of $p$--- then the bias of Laplacian smoothing can much smaller than that of kernel smoothing (for equivalent levels of variance).
+
+We emphasize that this does not contradict the well-known fact that kernel smoothing is rate-optimal for nonparametric regression.
+It simply reflects that in the standard nonparametric regression setup---which we adopt in the main part of this paper, and in which $P$ is assumed to be equivalent to Lebesgue measure--the biases of Laplacian eigenmaps and kernel smoothing are, up to constants, equivalent.
+There has been some work \textcolor{red}{(Wasserman, Niyogi, El Alaoui)} analyzing semi-supervised learning in the case where $p$ is very non-uniform, and $f_0$ ``adheres to'' $p$ in some way, such as the cluster assumption.
+However, a comprehensive analysis of the biases induced by Laplacian method as a function of $p$, how they differ with other approaches to estimation, and the consequences for nonparametric regression, remain outstanding.
 

### 2021-03-10 manuscripts/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
Properly differentiated between critical radius and squared critical radius in our text.

@@
 \end{equation*} 
-where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Clearly, the minimax critical radius $\epsilon$ depends on $\alpha$ and $b$.
+where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Clearly, the minimax critical radius $\epsilon(\mc{F})$ depends on $\alpha$ and $b$.
 However, we adopt the typical convention of treating $\alpha \in (0,1)$ and $b \geq 1$ as fixed positive constants; hence they will not affect the testing error rates, and we suppress them notationally.} See \citet{ingster82,ingster87,ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
 
-Testing $f_0 = 0$ is an easier problem than estimating $f_0$, and hence the minimax testing critical radius over $H^1(\Xset,M)$ is smaller than the minimax estimation rate, for $1 \leq d < 4$ (see \citet{ingster2009}):
+Testing $f_0 = 0$ is an easier problem than estimating $f_0$, and hence the minimax testing rate over $H^1(\Xset,M)$ is smaller than the minimax estimation rate.
+For $1 \leq d < 4$, the squared critical radius is (see \citet{ingster2009}):
 \begin{equation}
@@
 We show in the supplement that \smash{$\wh{f}$} is a level-$\alpha$ test.
-In the next theorem, we upper bound the worst-case risk \smash{$R_n(\wh{\varphi},H^1(\mc{X},M),\epsilon)$} of \smash{$\wh{\varphi}$}, whenever $\epsilon$ is at least (a constant times) the critical radius given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
+In the next theorem, we upper bound the worst-case risk \smash{$R_n(\wh{\varphi},H^1(\mc{X},M),\epsilon)$} of \smash{$\wh{\varphi}$}, whenever $\epsilon^2$ is at least (a constant times) the squared critical radius given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
 For this to hold, we will require a tighter range of scalings for the graph radius $r$.

### 2021-03-17 manuscripts/laplacian_eigenmaps/appendix.tex
Edited Section 2 Laplacian Eigenmaps manuscript.

@@
 
-In more detail, recall from our discuss.on in Section~\ref{subsec:analysis} that an essential step in our proofs is relating the iterated non-local operator $L_{P,\varepsilon}^j$ to the weighted Laplace-Beltrami operator $\Delta_P^j$, for $j = \floor{s/2}$.
-Suppose for the moment $s = 4$.
-Then we relate $L_{P,\varepsilon}$ to $\Delta_P$ by means of an intermediary nonlocal operator, as described by the following chain of estimates:
+\paragraph{Discussion of requirements that $s \leq 3$.}
+Recall from our discussion in Section~\ref{subsec:analysis} that an essential step in the proof of Lemma~\textcolor{red}{(?)}---which establishes a high-probability upper bound on $\dotp{L_{n,\varepsilon}^sf_0}{f_0}_n$---is relating the iterated non-local operator $L_{P,\varepsilon}^j$ to the weighted Laplace-Beltrami operator $\Delta_P^j$, for $j = \floor{s/2}$.
+Suppose for the moment $s = 3$, and $f \in C^3(\mc{X};1)$ and $p$ is uniform.
+In this case $\floor{s/2} = 1$, and the relation between $L_{P,\varepsilon}$ to $\Delta_P$ has been worked out by \textcolor{red}{(Burago2013, GarciaTrillos2019, Calder2019, GarciaTrillos2020)}.
+This relation is accomplished by means of an intermediary nonlocal operator, as described by the following chain of estimates:
 \begin{align*}
-	L_{P,\varepsilon} f(x) & = \int_{\mc{X}} \bigl(f(z) - f(x)\bigr)\eta\biggl(\frac{|z - x|}{\varepsilon}\biggr) \,dP(x) \\
-	& = \bigl(1 + O(\varepsilon^2)\bigr) \int_{\mc{X}} \bigl(f(z) - f(z)\bigr)\eta\biggl(\frac{d_{\mc{X}}(z,x)}{\varepsilon}\biggr) \,dP(x) \\
-	& = \Delta_Pf(x) + O\bigl(\varepsilon^2M\bigr) + O(\varepsilon^3).
+L_{P,\varepsilon} f(x) & = \frac{1}{\varepsilon^{m + 2}}\int_{\mc{X}} \bigl(f(z) - f(x)\bigr)\eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,d\mu(z) \\
+& \overset{(i)}{=} \frac{1}{\varepsilon^{m + 2}} \int_{\mc{X}} \bigl(f(z) - f(x)\bigr)\eta\biggl(\frac{d_{\mc{X}}(z,x)}{\varepsilon}\biggr) \,d\mu(z) + O(\varepsilon) \\
+& = \frac{1}{\varepsilon^{m + 2}} \int_{T_x(\mc{X})} \Bigl(f\bigl(\exp_x(v)\bigr) - f\bigl(\exp_x(0)\bigr)\Bigr)\eta\biggl(\frac{d_{\mc{X}}\bigl(\exp_x(v),\exp_x(0)\bigr)}{\varepsilon}\biggr) J_x(v) \,dv + O(\varepsilon) \\
+& \overset{(ii)}{=} \frac{1}{\varepsilon^{m + 2}} \int_{T_x(\mc{X})} \Bigl(f\bigl(\exp_x(v)\bigr) - f\bigl(\exp_x(0)\bigr)\Bigr)\eta\biggl(\frac{d_{\mc{X}}\bigl(\exp_x(v),\exp_x(0)\bigr)}{\varepsilon}\biggr) \,dv + O(\varepsilon) \\
+& = \frac{1}{\varepsilon^{m + 2}} \int_{T_x(\mc{X})} \Bigl(f\bigl(\exp_x(v)\bigr) - f\bigl(\exp_x(0)\bigr)\Bigr)\eta\biggl(\frac{\|v\|}{\varepsilon}\biggr) \,dv + O(\varepsilon) \\
+& \overset{(iii)}{=} \Delta_Pf(x) + O(\varepsilon)
 \end{align*}
-	Here, $d_{\mc{X}}(\cdot,\cdot)$ is the geodesic distance on $\mc{X}$.
-For simplicity we have assumed $f \in C^3(\mc{X})$ and $p \in C^2(\mc{X})$, but the main idea will not change if we assume Sobolev smoothness instead.
-From this, we may deduce that
+Here, $d_{\mc{X}}(\cdot,\cdot)$ is the geodesic distance on $\mc{X}$, $J_x(v)$ denotes the Jacobian of the exponential map $\exp_x$ evaluated at $v \in T_x(\mc{M})$, and $h = f \circ \exp_x$.
+Then $(i)$ follows because $\bigl|d_{\mc{X}}(x,z)  - \|z - x\|\bigr| = O(\varepsilon^3)$ and $\eta$ is a Lipschitz function with bounded Lipschitz constant, $(ii)$ follows because the Jacobian $J_x(v)$ satisfies $|J_x(v)| = 1 + O(\varepsilon^2)$ for all $v \in B(0,\varepsilon) \subset T_x(\mc{X})$, and $(iii)$ follows since the function $h$ admits the Taylor expansion
+\begin{equation*}
+h(v) = h(0) + \dotp{\nabla h(0)}{v} + \frac{1}{2} \dotp{\nabla^2 h(0) v}{v} + O(\varepsilon^3),~~\textrm{for all $v \in B(0,\varepsilon) \subset T_x(\mc{X})$.}
+\end{equation*}
+Then, using the upper bound on $\max_{i = 1,\ldots,n}|L_{n,\varepsilon}f(X_i) - L_{P,\varepsilon}(X_i)| = O_\Pbb((n\varepsilon^{m + 4})^{-1})$ given in \textcolor{red}{(?)}, and the fact that $\Delta_Pf \in C^1(\mc{X};1)$, we have that,
 \begin{align*}
-	L_{P,\varepsilon}^2 f(x) = 
+\dotp{L_{n,\varepsilon}^3f}{f}_n & = \frac{1}{n^2 \varepsilon^{m + 2}} \sum_{i,j = 1}^{n} \bigl(L_{n,\varepsilon}f(X_i) - L_{n,\varepsilon}f(X_j)\bigr)^2 \eta\biggl(\frac{\|X_i - X_j\|}{\varepsilon}\biggr) \\
+& = \frac{1}{n^2 \varepsilon^{m + 2}} \sum_{i,j = 1}^{n} \bigl(L_{P,\varepsilon}f(X_i) - L_{P,\varepsilon}f(X_j)\bigr)^2 \eta\biggl(\frac{\|X_i - X_j\|}{\varepsilon}\biggr) + O_{\Pbb}\biggl(\frac{1}{n\varepsilon^{m + 4}}\biggr) \\
+& = \frac{1}{n^2 \varepsilon^{m + 2}} \sum_{i,j = 1}^{n} \bigl(\Delta_Pf(X_i) - \Delta_Pf(X_j)\bigr)^2 \eta\biggl(\frac{\|X_i - X_j\|}{\varepsilon}\biggr) + O_{\Pbb}\biggl(\frac{1}{n\varepsilon^{m + 4}}\biggr) + O_{\Pbb}(1) \\
+& = O_{\Pbb}\biggl(\frac{1}{n\varepsilon^{m + 4}}\biggr) + O_{\Pbb}(1) = O_{\Pbb}(1),
 \end{align*}
-	When $s = 1$, this suffices.
+with the last equality following when $\varepsilon = \Omega(n^{-1/(m + 4)})$.
+
+On the other hand when $s = 4$, and thus $j = 2$, we have only that
+\begin{equation*}
+L_{P,\varepsilon}^2f(x) = L_{P,\varepsilon}(\Delta_Pf + O(\varepsilon)) = \Delta_P^2f(x) + O(1/\varepsilon),
+\end{equation*}
+which is useless when $\varepsilon \to 0$.
+The bottom line is that in the manifold case, the $O(\varepsilon^3)$ error incurred by going between Euclidean norm and geodesic distance is negligible only when $j = 1$.
+
+\section{Comparison of Laplacian eigenmaps and kernel smoothing in non-uniform density setting}
+
+\textcolor{red}{(TODO)}
+\begin{itemize}
+	\item \textcolor{red}{Mention the relationship between the estimation problem under consideration and classification with a margin (Tsybakov noise) condition.}
+\end{itemize}

### 2021-03-17 notes/weekly_notes/12.24.20.NotesForWeek.tex
Edited Section 2 Laplacian Eigenmaps manuscript.

@@
 \label{subsec:kernel_smoothing_bias_pf_1d_smooth_p}
-As in~\ref{subsec:kernel_smoothing_bias_pf_1d_uniform_p}, we take $\Omega = (0,1)$, let
+As in Section~\ref{subsec:kernel_smoothing_bias_pf_1d_uniform_p}, we take $\Omega = (0,1)$, let
 \begin{equation*}
-\|L_{P,\varepsilon}u\|_{P}^2 = \underbrace{\int_{\Omega_{\varepsilon}} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_1} + \underbrace{\int_{\partial_{\varepsilon}\Omega} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_2},
+\|L_{P,\varepsilon}u\|_{P}^2 = \underbrace{\int_{\Omega_{\varepsilon}} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_1} + \underbrace{\int_{\partial_{\varepsilon}\Omega} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_2}.
 \end{equation*}
-and turn to separately upper bounding $I_1$ and $I_2$.
+We upper bound $I_2$ immediately, using~\eqref{pf:kernel_smoothing_bias_1d_2}, the equivalent upper bound when $p$ is uniform, to determine that
+\begin{equation*}
+I_2 \leq p_{\max}\int_{\partial_{\varepsilon}\Omega} |L_{P,\varepsilon}u(x)|^2 \,dx \leq C\varepsilon^{2k + 1}\|u^{(k)}\|_P^2.
+\end{equation*}
+It therefore remains only to upper bound $I_1$.
 
 \paragraph{Upper bound on $I_1$.}
-We will show that for all $x \in \Omega_{\varepsilon}$,
+The $k$th order Taylor expansion of $u$ is
+\begin{equation}
+\label{pf:kernel_smoothing_bias_1d_smooth_p_1}
+u(z) - u(x) = \sum_{j = 1}^{k - 1}\frac{1}{j!} u^{(j)}(x) (z - x)^j + (z - x)^k r_z^{(k)}(x;u).
+\end{equation}
+The $k-1$st order Taylor expansion of $p$ is
+\begin{equation}
+\label{pf:kernel_smoothing_bias_1d_smooth_p_2}
+p(z) = \sum_{j = 0}^{k - 2}\frac{1}{j!} p^{(j)}(x) (z - x)^j + (z - x)^{k - 1} r_z^{(k - 1)}(x;p).
+\end{equation}
+Note that we adopt the convention that if $i > j$, then $\sum_{i}^{j} a_i = 0$, to cover the case when $k = 1$.
+Also, for a given $j \in \mathbb{N}$ and $j$ times differentiable function $f$, we have written $r_z^{(j)}(x;f)$ for the $j$th order remainder term,
 \begin{equation*}
-L_{P,\varepsilon}u(x) = \varepsilon^{k} \frac{\sigma_\eta}{k!}\Bigl( \frac{d^k}{dx^k}(up)(x) - p^{(k)}(x)u(x) + r(x)\Bigr)
+r_z^{(j)}(x;f) = \frac{1}{j!}\int_{0}^{1} (1 - t)^{j - 1} f^{(j)}(x + t(z - x)) \,dt.
 \end{equation*}
-for a remainder term $r$ satisfying $\int_{\Omega_{\varepsilon}} \bigl(r(x)\bigr)^2 p(x) \,dx  = o(1)$ as $\varepsilon \to 0$. This in turn will imply
+Now, we plug the expressions~\eqref{pf:kernel_smoothing_bias_1d_smooth_p_1} and~\eqref{pf:kernel_smoothing_bias_1d_smooth_p_2} in to the definition of $L_{P,\varepsilon}u(x)$ as follows (let $\ell = k - 1$ for simplicity):
+\begin{align}
+d_P(x) \cdot L_{P,\varepsilon}u(x) & = \int_{\Omega} (u(z) - u(x)) \eta\biggl(\frac{z - x}{\varepsilon}\biggr) p(z) \,dz \nonumber \\
+& = \sum_{j = 1}^{\ell} u^{(j)}(x) \int_{\Omega} (z - x)^j \eta\biggl(\frac{z - x}{\varepsilon}\biggr) p(z) \,dz + \int_{\Omega} (z - x)^{k} r_z^{(k)}(x;u) \eta\biggl(\frac{z - x}{\varepsilon}\biggr) p(z) \,dx \nonumber \\
+& = \sum_{j_1 = 1}^{\ell} \sum_{j_2 = 0}^{\ell - 1} \frac{u^{(j_1)}(x) p^{(j_2)}(x)}{j_1!j_2!} \int_{\Omega} (z - x)^{j_1 + j_2} \eta\biggl(\frac{z - x}{\varepsilon}\biggr) p(z) \,dz + \sum_{j = 1}^{\ell} u^{(j)}(x) \int_{\Omega} (z - x)^{j + \ell} r_z^{(\ell)}(x;p) \,dz \nonumber \\
+& ~~ + \int_{\Omega} (z - x)^{k} r_z^{(k)}(x;u) \eta\biggl(\frac{z - x}{\varepsilon}\biggr) p(z) \nonumber \\
+& =: T_1(x) + T_2(x) + T_3(x). \label{pf:kernel_smoothing_bias_1d_smooth_p_3}
+\end{align}
+
+By Assumptions~\ref{asmp:domain} and~\ref{asmp:density}, we have that $|d_P(x)| \geq p_{\min} \alpha\varepsilon$ for all $x \in \Omega$. It will therefore suffice to show the following upper bounds
+\begin{equation}
+\label{pf:kernel_smoothing_bias_1d_smooth_p_4}
+\|T_i\|_{\Omega_{\varepsilon}}^2 \leq C_i \varepsilon^{2(k + 1)} \|u\|_{H^{k}(P)}^2
+\end{equation}
+for $i = 1,2,3$, since along with~\eqref{pf:kernel_smoothing_bias_1d_smooth_p_3} this will imply
+\begin{equation*}
+I_1 \leq \frac{3p_{\max}}{\alpha^2 p_{\min}^2 \varepsilon^2} \sum_{i = 1}^{3} C_i \varepsilon^{2(k + 1)} \|u\|_{H^{k}(P)}^2 = C \varepsilon^{2k} \|u\|_{H^k(P)}^2.
+\end{equation*}
+We now establish~\eqref{pf:kernel_smoothing_bias_1d_smooth_p_4} for each of $i = 1,2,3$.
+
+\textit{Upper bound on $\|T_1\|_{\Omega_{\varepsilon}}^2$.}
+We introduce some notation to keep our formulae as readable as possible.
+Let $[k] = \{1,\ldots,k\}$ and let $\mathbf{J} := \bigl\{(j_1,j_2): j_1 \in [\ell], j_2 \in [\ell - 1] \cup \{0\}, j_1 + j_2 \geq k\bigr\}$.
+Note that if $j_1 \in [\ell]$ and $j_2 \in [\ell - 1] \cup \{0\}$ do not satisfy $(j_1,j_2) \in \mathbf{J}$, then the order-$s$ property of $\eta$ combined with the fact $B(x,\varepsilon) \in \Omega$ implies that
+\begin{equation*}
+\int_{\Omega} (z - x)^{j_1 + j_2} \eta\biggl(\frac{z - x}{\varepsilon}\biggr) p(z) \,dz = 0.
+\end{equation*}
+On the other hand if $(j_1,j_2) \in \mathbf{J}$ then by a change of variables
+\begin{equation*}
+\int_{\Omega} (z - x)^{j_1 + j_2} \eta\biggl(\frac{z - x}{\varepsilon}\biggr) p(z) \,dz = \varepsilon^{j_1 + j_2} \int_{B(0,1)} v^{j_1 + j_2 + 1} \eta(|v|) \,dv =: \varepsilon^{j_1 + j_2 + 1} \sigma_\eta^{(j_1 + j_2)},
+\end{equation*}
+with Assumption~\ref{asmp:kernel} guaranteeing that $\sigma_\eta^{(j)} < \infty$ for all $j \in \mathbb{N}$. Therefore,
+\begin{equation*}
+T_1(x) = \sum_{(j_1,j_2) \in \mathbf{J}} \frac{\varepsilon^{j_1 + j_2 + 1}}{j_1!j_2!} u^{(j_1)}(x) p^{(j_2)}(x) .
+\end{equation*}
+Next, we note that by H\"{o}lder's inequality, $\|u^{(j_1)} p^{(j_2)} \|_{\Omega_{\varepsilon}}^2 \leq  \|p^{(j_2)}\|_{L^{\infty}(\Omega_{\varepsilon})}^2 \|u^{(j_1)}\|_{\Omega_{\varepsilon}}^2 \leq \|u\|_{H^k(\Omega)}^2$.
+As a result,
+\begin{equation*}
+\|T_1\|_{\Omega_{\varepsilon}}^2 \leq |\mathbf{J}| \cdot \sum_{(j_1,j_2) \in \mathbf{J}} \frac{\varepsilon^{2(j_1 + j_2 + 1)}}{(j_1!j_2!)^2} \|p^{(j_1)}  u^{(j_2)}\|_{\Omega_{\varepsilon}}^2 \leq \frac{1}{(k - 1)!} \varepsilon^{2(k + 1)} \|u\|_{H^k(\Omega)}^2,
+\end{equation*}
+and thus~\eqref{pf:kernel_smoothing_bias_1d_smooth_p_3} follows with $C_1 = 1/(k - 1)!$.
+
+\textit{Upper bound on $\|T_2\|_{\Omega_{\varepsilon}}^2$.}
+Note that if $k = 1$ then $T_2(x) = 0$, so that we need only cover the case where $k > 1$.
+In this case, observe that since $p \in C^{\ell}(P;1)$, it follows that $|r_z^{(\ell)}(x;p)| \leq 1/\ell!$ for all $z \in B(x,\varepsilon)$, and thus by H\"{o}lder's inequality,
+\begin{equation*}
+\biggl|\int_{\Omega} (z - x)^{j + \ell} r_z^{(\ell)}(x;p) \eta\biggl(\frac{z - x}{\varepsilon}\biggr) \,dz\biggr| \leq \varepsilon^{j + \ell}\int_{\Omega} \biggl| \eta\biggl(\frac{z - x}{\varepsilon}\biggr) \biggr| \,dz \leq \|\eta\|_{\infty} \nu_1 \varepsilon^{j + \ell + 1}
+\end{equation*}
+for any $j \in \mathbb{N}$. An upper bound on $|T_2(x)|$ follows immediately,
+\begin{equation*}
+|T_2(x)| \leq \sum_{j = 1}^{\ell} \biggl|\frac{u^{(j)}(x)}{j!}\biggr| \biggl|\int_{\Omega} (z - x)^{j + \ell} r_z^{(\ell)}(x;p) \eta\biggl(\frac{z - x}{\varepsilon}\biggr) \,dz\biggr| \leq \|\eta\|_{\infty} \nu_1 \varepsilon^{k + 1} \sum_{j = 1}^{\ell} \frac{|u^{(j)}(x)|}{j!\ell!}.
+\end{equation*}
+and calculations similar to above imply that
+\begin{equation*}
+\|T_2\|_{\Omega_{\varepsilon}}^2 \leq \biggl(\frac{\|\eta\|_{\infty} \nu_1}{(\ell - 1)!}\biggr)^2 \varepsilon^{2(k + 1)} \|u\|_{H^k(\Omega)}^2,
+\end{equation*}
+and thus~\eqref{pf:kernel_smoothing_bias_1d_smooth_p_3} follows with $C_2 = \|\eta\|_{\infty} \nu_1/(\ell - 1)!$.
+
+\textit{Upper bound on $\|T_3\|_{\Omega_{\varepsilon}}^2$.}
+We begin by upper bounding $|T_3(x)|$ for any $x \in \Omega_{\varepsilon}$. Using the change of variables $v = (z - x)/\varepsilon$, we determine that
+\begin{equation*}
+|T_3(x)| \leq \int_{\Omega} \biggl| (z - x)^{k} r_z^{(k)}(x;u) \eta\biggl(\frac{z - x}{\varepsilon}\biggr)  p(z) \biggr|\,dz  \leq \varepsilon^{k + 1} \|\eta\|_{\infty} p_{\max} \int_{B(0,1)} \Bigl|r_{x + \varepsilon v}^{(k)}(x;u)\Bigr| \,dv.
+\end{equation*}
+From here, using the Cauchy-Schwarz inequality twice, we obtain the following upper bound on $\|T_3\|_{\Omega_{\varepsilon}}^2$,
 \begin{align*}
-I_1 = \int_{\Omega_{\varepsilon}} |L_{P,\varepsilon}u(x)|^2 \,dx & = \varepsilon^{2k} \frac{\sigma_\eta^2}{(k!)^2}\Bigl(\int_{\Omega_{\varepsilon}} \Bigl( \frac{d^k}{dx^k}(up)(x) - p^{(k)}(x)u(x) + r(x)\Bigr)^2  p(x) \,dx \\
-& \leq 3 \varepsilon^{2k} \frac{\sigma_\eta^2}{(k!)^2} \bigl( |up|_{H^k(P)}^2 + |p^{(k)}u|_{\Leb^2(P)}^2 + o(1)\bigr) \\
-& \leq C \varepsilon^{2k} \frac{\sigma_\eta^2}{(k!)^2} \Bigl( \|p\|_{C^k(P)}^2\bigl(|u|_{H^k(P)}^2 + |u|_{\Leb^2(P)}^2\bigr) + o(1) \Bigr),
+\|T_3\|_{\Omega_{\varepsilon}}^2 & \leq (\|\eta\|_{\infty} p_{\max})^2 \varepsilon^{2(k + 1)} \int_{\Omega_{\varepsilon}}\biggl(\int_{B(0,1)} \Bigl|r_{x + \varepsilon v}^{(k)}(x;u)\Bigr| \,dv\biggr)^2 \,dx \\
+& \leq (\|\eta\|_{\infty} p_{\max} \nu_1)^2 \varepsilon^{2(k + 1)} \int_{\Omega_{\varepsilon}} \int_{B(0,1)}\Bigl|r_{x + \varepsilon v}^{(k)}(x;u)\Bigr|^2 \,dv \,dx \\
+& \leq \biggl(\frac{\|\eta\|_{\infty} p_{\max} \nu_1}{(k - 1)!}\biggr)^2 \varepsilon^{2(k + 1)} \int_{\Omega_{\varepsilon}} \int_{B(0,1)} \int_{0}^{1} \bigl|u^{(k)}\bigl(x + t\varepsilon v\bigr)\bigr|^2 \,dv \,dx.
 \end{align*}
-with the last line following from Young's inequality.
+Finally, exchanging the order of integration gives
+\begin{equation*}
+\|T_3\|_{\Omega_{\varepsilon}}^2 \leq \biggl(\frac{\|\eta\|_{\infty} p_{\max}}{(k - 1)!}\biggr)^2 \nu_1 \varepsilon^{2(k + 1)} \|u\|_{H^k(\Omega)}^2,
+\end{equation*}
+and thus~\eqref{pf:kernel_smoothing_bias_1d_smooth_p_3} follows with $C_3 = (\|\eta\|_{\infty} p_{\max}/(k - 1)!)^2 \nu_1$.
 
-It remains to show the pointwise estimate \textcolor{red}{(...)}.
+\textcolor{red}{(TODO)}
+\begin{itemize}
+	\item \textcolor{red}{Equation~\eqref{pf:kernel_smoothing_bias_1d_smooth_p_3} is not well formatted}.
+	\item \textcolor{red}{Need to define the norm used on the left hand side of  Equation~\eqref{pf:kernel_smoothing_bias_1d_smooth_p_4}}.
+\end{itemize}
+
+\section{Proof of Lemma~\ref{lem:kernel_smoothing_bias}}
+Letting $\Omega_{\varepsilon} = \{x: B(x,\varepsilon) \subset \Omega\}$ and $\partial_{\varepsilon}\Omega = \Omega \setminus \Omega_{\varepsilon}$, we have that 
+\begin{equation*}
+\|L_{P,\varepsilon}u\|_P^2 = \underbrace{\int_{\Omega_{\varepsilon}} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_1} + \underbrace{\int_{\partial_{\varepsilon}\Omega} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_2}
+\end{equation*}
+We will separately upper bound $I_1$---using the higher order property of $\eta$---and $I_2$---using the zero-trace property of $u$ and its $1$st up through $j - 1$st order derivatives.
+
+\paragraph{Upper bound on $I_1$.}
+
+\paragraph{Upper bound on $I_2$.}
 

### 2021-03-17 manuscripts/laplacian_eigenmaps/appendix.tex

@@
 \noindent
+
+\section{Graph-dependent error bounds}
+\label{sec:fixed_graph_error_bounds}
+In this section, we adopt the fixed design perspective; or equivalently, condition on $X_i = x_i$ for $i = 1,\ldots,n$.
+Let $G = \bigl([n],W\bigr)$ be a fixed graph on $\{1,\ldots,n\}$ with Laplacian matrix $L = \sum_{k = 1}^{n}\lambda_k v_k v_k^{\top}$.
+The randomness thus all comes from the responses
+\begin{equation}
+\label{eqn:fixed_graph_regression_model}
+Y_i = f_{0}(x_i) + \varepsilon_i
+\end{equation}
+where the noise variables $\varepsilon_i$ are independent $N(0,1)$.
+In the rest of this section, we will mildly abuse notation and write $f_0 = (f_0(x_1),\ldots,f_0(x_n)) \in \Reals^n$.
+We will also write ${\bf Y} = (Y_1,\ldots,Y_n)$.
+
+Let $\wh{T} = \sum_{k = 1}^{K} \dotp{{\bf Y}}{v_k}_n^2$, and let $\varphi = \1\{\wh{T} \geq t_a\}$. In the following Lemma, we upper bound the Type I and Type II error of the test $\varphi$.
+
+\begin{lemma}
+	Suppose we observe $(Y_1,x_1),\ldots,(Y_n,x_n)$ according to~\eqref{eqn:fixed_graph_regression_model}.
+	\begin{itemize}
+		\item If $f_0 = 0$, then $\Ebb_0[\varphi] \leq a$.
+		\item Suppose $f_0 \neq 0$ satisfies
+		\begin{equation*}
+		\sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 \geq 16\frac{\sqrt{2K}}{n}\biggl(\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr).
+		\end{equation*}
+		Then $\Ebb_{f_0}[1 - \phi] \leq b$.
+	\end{itemize}
+\end{lemma}
 

### 2021-03-17 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex

--- 2021-03-08 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
@@
 \usepackage{tikz}
+\usepackage{bm}
+\usepackage{multirow}
 
 \usepackage{natbib}
-\renewcommand{\bibname}{REFERENCES}
+\renewcommand{\bibname}{References}
 \renewcommand{\bibsection}{\subsubsection*{\bibname}}
@@
 \newcommand{\Rd}{\Reals^d}
+\newcommand{\RD}{\Reals^D}
 \newcommand{\Nbb}{\mathbb{N}}
@@
 
-\textbf{(1) Geometric graphs.} In graph-based learning, one observes data $X_1,\ldots,X_n$ sampled independently from an unknown distribution $P$, and forms a geometric graph $G$---with edges $e_{ij}$ corresponding to proximity between samples $X_i$ and $X_j$---over the observed data.
-Geometric graphs encode information about $P$ in an extremely general manner, and can be leveraged to conduct many different fundamental statistical tasks.
-These include clustering, semi-supervised learning, classification, and regression---both estimation and goodness-of-fit testing.
-Though much theoretical work has been done on the consistency of graph-based learning methods---and more recently, rates of convergence have been established for some problems---little is known so far about their optimality, even for classic statistical tasks.
-
-\textbf{(2) What we study: Regression and Laplacian eigenmaps.} In this paper we focus on regression where in addition to the design points $X_1,\ldots,X_n$ one observes real-valued responses $Y_1,\ldots,Y_n$, and seeks to learn the unknown regression function $f_0(x) := E[Y|X = x]$.
-We consider both the estimation and goodness-of-fit testing problems.
-The methods we will study are based on \emph{Laplacian eigenmaps}, first introduced by \cite{belkin03a}, which projects the response vector $Y = (Y_1,\ldots,Y_n)$ onto the span of the leading eigenvectors of the graph Laplacian $L$.
+Suppose we observe data $X_1,\ldots,X_n$ sampled independently from an unknown distribution $P$.
+As a replacement for $P$, we form a geometric graph $G$ over the observed data, with weighted edges $W_{ij}$ corresponding to proximity between samples $X_i$ and $X_j$.
+Geometric graphs encode information about $P$--- for instance, the number of connected components of its support, and the shape of each connected component---in an extremely general manner, and can be leveraged to conduct many different fundamental statistical tasks.
+These include clustering, manifold learning, semi-supervised learning, classification, and regression---both estimation and goodness-of-fit testing.
+Though much theoretical work has been done on the consistency of learning methods defined with respect to geometric graphs---and more recently, rates of convergence have been established for some methods---little is known so far about their optimality, even for classic statistical tasks.
+
+In this paper we focus on the theoretical statistical properties of \textcolor{red}{graph-based regression}.
+We assume that in addition to the design points $X_1,\ldots,X_n$ one observes real-valued responses $Y_1,\ldots,Y_n$, and seeks to learn the unknown regression function $f_0(x) := E[Y|X = x]$.
+The methods we will study are based on \emph{Laplacian eigenmaps}, first introduced by \cite{belkin03a}, which takes the eigenvectors of the graph Laplacian $L$.
 The Laplacian $L$ is a difference operator, acting on functions $f: \{X_1,\ldots,X_n\} \to \Reals$ as follows,
@@
 \label{eqn:graph_laplacian}
-Lf(X_i) = \sum_{j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)e_{ij}. 
+Lf(X_i) = \sum_{j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)W_{ij}. 
 \end{equation}
@@
 
-\textbf{(3) Why we study.} Laplacian eigenmaps is thus an example of a \emph{spectral series} estimator.
+Laplacian eigenmaps is thus an example of a \emph{spectral series} estimator.
 A spectral series estimator is a special case of an orthogonal series estimator, one of the most classical methods of non-parametric regression.
-Traditionally, orthogonal series estimators are defined with respect to a fixed set of basis functions, orthogonalized with respect to some pre-determined reference measure $Q$.
+Traditionally, orthogonal series estimators are defined with respect to a fixed set of basis functions, orthogonalized in $L^2(Q)$ with respect to some pre-determined reference measure $Q$.
 Spectral series estimators take a particular orthogonal basis, the eigenfunctions of $\Delta_Q$.
@@
 
-The theoretical properties of (classical) spectral series estimators are by this point well-understood \textcolor{red}{(references)}.
-In particular, they are minimax optimal for nonparametric regression over H\"{o}lder and Sobolev spaces.
-However, in practice these estimators suffer from some serious drawbacks.
-Finding the eigenfunctions of $\Delta_Q$ is in general non-trivial.
+The theoretical properties of (classical) spectral series methods are various, and are in general extremely well-understood \textcolor{red}{(references)}.
+In particular, they are minimax rate optimal for nonparametric regression over H\"{o}lder and Sobolev spaces.
+However, in practice these methods suffer from some serious drawbacks.
+As a practical matter,  even finding the eigenfunctions of $\Delta_Q$ is in general non-trivial.
 The statistical optimality of such estimators also hinges on the reference measure $Q$ being equal to $P$, an unrealistic assumption when $P$ is unknown.
-If $Q \neq P$, the standard recommendation is to instead do least squares---this can be statistically rate-optimal, but is numerically unstable when dimension $d$ is even moderately large.
-
-\textbf{(4) Our contributions.} Laplacian eigenmaps avoid these drawbacks \textcolor{red}{(...)} On the other hand their statistical properties are not as well understood \textcolor{red}{(...)}.
-The primary contribution of our paper is to fill this theoretical gap, by answering the following question:
+If $Q \neq P$, the standard recommendation is to instead perform nonparametric least squares---this can be statistically rate-optimal, but is numerically unstable when the dimension $d$ is even moderately large.
+
+\textcolor{red}{(TODO)}: Change the preceding paragraph.
+
+Laplacian eigenmaps avoids these drawbacks \textcolor{red}{(...)} On the other hand its statistical properties are not as well understood \textcolor{red}{(...)}. Thus, there exists a theoretical gap.
+
+\paragraph{Our contributions.} The primary contribution of our paper is to fill this theoretical gap, by answering the following question:
 
@@
 
-Broadly speaking, we show that when the regression function $f_0$ is smooth, in the sense of having bounded derivatives in the Sobolev sense, Laplacian eigenmaps is statistically minimax optimal.
+Broadly speaking, we show that when the regression function $f_0$ is smooth, in the Sobolev sense of having weak derivatives bounded in $\Leb^2$-norm, Laplacian eigenmaps methods are statistically minimax optimal, for both estimation and goodness-of-fit testing.
 This statement holds for different relations between the dimension $d$ and number of derivatives $s$, depending on the problem (estimation or testing).
-\textcolor{red}{(TODO): Add two tables summarizing your major results here.}
-
-\textbf{(5) Organization.}
-
-\textcolor{red}{(TODO)}
-
-\textbf{(6) Notation.}
-
-\textcolor{red}{TODO}: Define $\Leb^2(\mc{X}), \Leb^2(P)$ and $\Leb^2(P_n)$.
-
-For a square, symmetric matrix $A \in \Reals^{n \times n}$, an eigenvector $v$ of $A$ is a \textcolor{red}{unit-norm} solution to the eigenproblem
-\begin{equation*}
-Av = \lambda v,
-\end{equation*}
-and $\lambda$ is the corresponding eigenvalue.
-
-\section{Setup, Background, and Overview of Results}
-\label{sec:setup_main_results}
-
-In this section, we begin by giving a precise definition of Laplacian eigenmaps.
-We then review minimax rates for non-parametric regression over Sobolev spaces, allowing us to summarize our main results.
-
-\subsection{Non-parametric regression with Laplacian Eigenmaps}
-\label{sec:regression_laplacian_eigenmaps}
-
-\textbf{(1) Formal setup.}
-We will operate in the usual setting of non-parametric regression with random design.
-We observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$, where the design points $X_1,\ldots,X_N$ are drawn i.i.d from a distribution $P$ supported on a compact set $\mc{X} \subseteq \Rd$, and responses
-\begin{equation}
-\label{eqn:model}
-Y_i = f_0(X_i) + \epsilon_i,
-\end{equation}
-with regression function $f_0: \mc{X} \to \Reals$, and $\epsilon_i \sim N(0,1)$ independent Gaussian noise.
-For simplicity we will assume throughout that the noise has unit-variance, but all of our results extend in a straightforward manner to the case where the variance is equal to a known positive value.
-
-\textbf{(2) Laplacian eigenmaps.} Laplacian eigenmaps constructs an estimate of the regression function $f_0$ using the eigenvectors of a neighborhood graph Laplacian operator.
-For a positive, radially symmetric kernel $\eta: [0,\infty) \to [0,\infty)$, and a bandwidth parameter $\varepsilon > 0$, the \emph{neighborhood graph Laplacian} operator $L_{n,\varepsilon}$ is
-\begin{equation}
-\label{eqn:neighborhood_graph_laplacian}
-L_{n,\varepsilon}u(x) = \frac{1}{n\varepsilon^{d + 2}} \sum_{i = 1}^{n} \bigl(u(x) - u(X_j)\bigr) \eta\biggl(\frac{\|x - X_j\|}{\varepsilon}\biggr)
-\end{equation}
-The constant scaling $(n\varepsilon^{d + 2})^{-1}$ is purely for convenience in taking limits as $n \to \infty, \varepsilon \to 0$.
-Also, although~\eqref{eqn:neighborhood_graph_laplacian} makes sense for any $x \in \mc{X}$ and $u: \mc{X} \to \Reals$, we will typically think of $x \in \{X_1,\ldots,X_n\}$ and $u = (u(X_1),\ldots,u(X_n)) \in \Leb^2({\bf X})$.
-The graph Laplacian is positive semi-definite, and we will index its eigenpairs $(\lambda_1,v_1),\ldots,(\lambda_n,v_n)$ in ascending order of eigenvalue, $0 = \lambda_1 \leq \ldots \leq \lambda_n$.
-
-For a given $K \in \{1,\ldots,n\}$, the order-$K$ Laplacian eigenmaps estimator simply projects the response vector ${\bf Y}$ onto the first $K$ eigenvectors of $L_{n,\varepsilon}$: letting $V_K \in \Reals^{n \times K}$ be the matrix with columns $v_1,\ldots,v_K$, we have that
-\begin{equation}
-\label{eqn:laplacian_eigenmaps_estimator}
-\wh{f} := \sum_{k = 1}^{K} \dotp{Y}{v_k}_{n} v_k = \frac{1}{n} V_K V_K^{\top} Y.
-\end{equation} 
-Since the eigenvectors $v_1,\ldots,v_n$ are orthogonalized with respect to the $L^2(P_n)$ inner product, the estimator $\wh{f}$ is the least-squares solution to the linear regression problem with responses $Y_1,\ldots,Y_n$ and features $v_1,\ldots,v_K$.
-Note that $\wh{f}$ is defined only in-sample, that is, only at the design points $X_1,\ldots,X_n$---we will consider the question of out-of-sample estimation later in Section~\ref{sec:out_of_sample}.
-
-If $\wh{f}$ is a reasonable estimate of $f_0$, then the test statistic
-\begin{equation}
-\label{eqn:laplacian_eigenmaps_test}
-\wh{T} := \|\wh{f}\|_n^2
-\end{equation}
-is in turn a reasonable estimate of $\|f_0\|_{P}^2$, and can be used to distinguish whether or not $f_0 = 0$.
-
-\subsection{Sobolev Classes}
-\label{sec:sobolev}
-Before we give an overview of our main results, we will first pause to review the Sobolev classes and \textcolor{red}{associated} minimax rates.
-
-\textbf{(1) Sobolev norms, semi-norms, and balls.} 
-The Sobolev class $H^s(P)$ consists of all those functions $f \in \Leb^2(P)$ which have weak partial derivatives of all orders $j = 1,\ldots,s$, all of which are bounded in $\Leb^2(P)$ norm.
-The Sobolev norm $\|f\|_{H^s(P)}$ is defined according to
-\begin{equation*}
-\|f\|_{H^s(P)}^2 := \|f\|_{\Leb^2(P)}^2 + \sum_{j = 1}^{s} \|D^jf\|_{\Leb^2(P)}^2,
-\end{equation*}
-where $D^jf := \sum_{|\alpha| = j}D^{\alpha}f$ is the sum of all weak partial derivatives of order $j$, and we use the multiindex notation $D^{\alpha}f = \partial^j f/(\partial^{\alpha_1}x^1\ldots \partial^{\alpha_d}x^d)$ for multiindex $\alpha$.
-The Sobolev ball of radius $M$ consists of all those functions $f \in H^s(P)$ with Sobolev norm no greater than $M$,
-\begin{equation*}
-H^s(P;M) := \bigl\{f \in H^s(P): \|f\|_{H^s(P)} \leq M\bigr\}
-\end{equation*}
-
-\textcolor{red}{(TODO)}: 
-\begin{itemize}
-	\item Possibly define the iterated Laplace-Beltrami operator.
-Point out that for $H^{2s}(P)$ functions, the Sobolev semi-norm can be stated in terms of inner products using the iterated Laplace-Beltrami operator.
-	\item Introduce the notation $H^s(\mc{X}) = H^s(\nu)$, where $\nu$ is Lebesgue measure.
-\end{itemize}
-
-\textbf{(2) Zero-trace and periodic conditions.}
-When $s \geq 1$, our results regarding Laplacian eigenmaps will require that $f_0$ satisfy some boundary assumptions.
-In particular, we will assume that $f_0$ belongs to the set of zero trace Sobolev functions $H_0^{s}(P)$, meaning \textcolor{red}{roughly} that $f_0 \in H^s(P)$ additionally satisfies,
-\begin{equation}
-\label{eqn:zero_trace}
-D^jf(x) = 0,~~\textrm{for all $x \in \partial \mc{X}$ and $j = 0,\ldots,s - 1$.}
-\end{equation}
-\textcolor{red}{(TODO)}: This is in spirit right but technically incoherent unless $f \in C^{\infty}(\mc{X})$. Need to use the formal definition of trace.
-
-Let us justify why we impose a boundary condition such as zero-trace.
-For this part only we concentrate on the special case of $d = 1$ and the domain $\mc{X} = [0,1]$.
-In this case, there exists another way to define Sobolev space, via a sequence space representation (equiv Fourier transform).
-Let $\psi_1,\psi_2,\ldots$ denote the trigonometric basis of $\Leb^2([0,1])$,\footnote{To be explicit, $\psi_1(x) := 1$, $\psi_{2k}(x) := \sqrt{2} \cos(2\pi k x)$ and $\psi_{2k + 1}(x) := \sqrt{2} \sin(2\pi k x)$} and let $\Theta \subset \ell^2(\mathbb{N})$ be the Sobolev ellipsoid $\Theta = \{\theta \in \ell^2(\mathbb{N}): \sum_{k = 1}^{\infty} k^{2s} \theta_k^2 \leq M\}$.
-The function class $\{f = \sum_{k = 1}^{\infty} \theta_k \psi_k: \theta \in \Theta\}$ is \textcolor{red}{equivalent to} the \emph{periodic Sobolev space} $H_{\mathrm{per}}^s([0,1])$, defined as
-\begin{equation}
-\label{eqn:periodic_sobolev_space}
-H_{\mathrm{per}}^s([0,1]) = \Bigl\{f \in H^s([0,1],M): D^jf(0) = D^jf(1)~~\textrm{for $j = 0,\ldots, s - 1$} \Bigr\}
-\end{equation}
-It is known that when $P$ is uniform, the eigenvectors $v_k$ converge to $\psi_k$ as $n \to \infty, \varepsilon \to 0$, and it is therefore natural that they accurately approximate only periodic functions $f_0 \in H^s([0,1])$.
-Although the zero-trace boundary condition is somewhat more restrictive than mere periodicity, the point is that some kind of boundary condition is inherently necessary in order to obtain sharp rates for spectral series estimators.
-For simplicity, we will stick to the specific boundary condition~\eqref{eqn:zero_trace}.
-
-\subsection{Minimax Rates}
-\label{subsec:minimax_rates_sobolev}
-We now turn to reviewing the minimax estimation and goodness-of-fit testing rates over Sobolev classes.
-We will always be interested in measuring loss in mean-squared error, that is in squared $\Leb^2$-norm.
-In this case, standard non-parametric minimax rates hold under some regularity conditions on $\mc{X}$ and $P$.
-Specifically, we will assume the following.\footnote{Assumption~\ref{asmp:sobolev_radius} ensures that we get a non-parametric rate.
-If $M \lesssim n^{-1/2}$ then both the estimation and testing rates become the parametric $n^{-1}$.}
-\begin{enumerate}[label=(A\arabic*)]
-	\item 
-	\label{asmp:domain}
-	The domain $\mc{X}$ is an open connected subset of $\Rd$, with Lipschitz boundary.
-	\item
-	\label{asmp:density}
-	The design distribution $P$ admits a density $p$ with respect to Lebesgue measure. The density $p$ is bounded away from $0$ and $\infty$: there exists $\rho \geq 1$ such that
-	\begin{equation*}
-	0 < \frac{1}{\rho} \leq p(x) \leq \rho < \infty,~~\textrm{for all $x \in \mc{X}$}
-	\end{equation*}
-	\item
-	\label{asmp:sobolev_radius}
-	The radius $M$ of the Sobolev class satisfies $M = M(n) \gtrsim n^{-1/2}$.
-\end{enumerate}
-Under assumptions~\ref{asmp:domain}-\ref{asmp:sobolev_radius}, the minimax estimation rate over $\mc{H} = H^s(\mc{X};M)$ or $\mc{H} = H_0^s(\mc{X};M)$) is (see e.g. \citet{tsybakov2008_book})
-\begin{equation}
-\label{eqn:minimax_estimation_rate}
-\inf_{\wh{f}} \sup_{f_0 \in \mc{H}} \|f_0 - f\|_P^2 \asymp M^{2d/(2s + d)}n^{-2s/(2s + d)};
-\end{equation}
-here the infimum is over estimators $\wh{f}$.
-
-In the goodness-of-fit testing problem, we ask for a test function---formally, a Borel measurable function $\phi$ that takes values in $\{0,1\}$--- which can distinguish between the hypotheses
-\begin{equation}
-\mathbf{H}_0: f_0 = f_0^{\star}, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{H} \setminus \{f_0^{\star}\}.
-\end{equation} 
-Typically, the null hypothesis $f_0 = f_0^{\star} \in \mc{F}$ reflects the absence of interesting structure, and $\mc{F} \setminus  \{f_0^{\star}\}$ is a set of smooth departures from this null.
-To fix ideas, as in \citet{ingster2009} we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\star} = 0$ and $\mc{H}$ is a Sobolev ball, $\mc{H} = H^s(\Xset,M)$ or $\mc{H} = H_0^s(\mc{X};M)$ is a Sobolev ball; this is without loss of generality.
-
-The minimax critical radius is the smallest value of $\sigma$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{H}_{\sigma} := \mc{H} \cap \{f: \|f\|_{\Leb^2(P)} \geq \sigma\}$:
-\begin{equation*}
-\sigma(\mc{H}) := \inf\Biggl\{\sigma > 0: \inf_{\phi} \biggl[ \sup_{f_0 \in \mc{H}_{\sigma}} \Ebb_{f_0}[1 - \phi]\biggr] \leq \alpha\Biggr\}
-\end{equation*} 
-where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation under the regression function $f_0$.
-Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing critical radius over $\mc{H}$ is much smaller than the minimax estimation rate \citep{ingster2009}:
-\begin{equation}
-\label{eqn:sobolev_space_testing_critical_radius}
-\sigma^2(\mc{H}) \asymp M^{2d/(4s + d)}n^{-4s/(4s + d)}~~\textrm{for $1 \leq d < 4s$.}
-\end{equation}
-When $4s \geq d$ the functions in $H^s(\Xset)$ are very irregular---formally speaking $H^s(\Xset)$ does not \textcolor{red}{continuously} embed into $\Leb^4(\Xset)$ when $4s \geq d$---and the minimax testing rates in this regime are unknown.
-
-\textcolor{red}{(TODO)}
-\begin{itemize}
-	\item Is there a better citation than \citet{tsybakov2008_book} for the minimax rate of estimation~\eqref{eqn:minimax_estimation_rate}?
-Tsybakov considers the univariate case with fixed, equispaced design.
-	\item I copy pasted the testing part directly from our AISTATS paper, then made a few necessary changes. Is that bad?
-	\item I have stated the rate in~\eqref{eqn:sobolev_space_testing_critical_radius} over $\mc{H}$.
-In reality,~\citep{ingster2009} make some kind of boundary assumptions and I cannot remember exactly what.
-Is it ok to state it as is, or do I need to be more precise?
-\end{itemize}
-
-\subsection{Spectral series methods for Sobolev Spaces}
-Spectral series estimators---and tests using their $\Leb^2$-norm---are particularly designed for regression in Sobolev spaces.
-Let $f_1,f_2,\ldots$ be the unit norm eigenfunctions of $\Delta_P$.
-The classical order-$K$ spectral series estimator $\wt{f}$ and test statistic $\wt{T}$ are
-\begin{equation*}
-\wt{f} = \sum_{k = 1}^{K} \dotp{Y}{f_k}_n f_k~~\textrm{and}~~\wt{T} = \|\wt{f}\|_P^2.
-\end{equation*}
-The estimator $\wt{f}$ and a test based on $\wt{T}$ are minimax rate-optimal over $H_0^s(\mc{X})$.
-This fact is intimately related to the series representation of periodic Sobolev spaces, e.g.~\eqref{eqn:periodic_sobolev_space}, whereby the model~\eqref{eqn:model} can be formally reduced to the Gaussian sequence model (see e.g.
-\citep{tsybakov08,gine16}).
-
-Unfortunately, in practice the estimator $\wt{f}$ is usually unworkable.
-Typically one does not have direct knowledge of the design distribution $P$, and even if $P$ is known, the eigenfunctions of $\Delta_P$ are rarely easy to compute.
-As an alternative, suppose instead we had access only to eigenfunctions $f_1',f_2',\ldots$ of the unweighted Laplace-Beltrami operator $\Delta_{\nu}$.
-Then letting $F_K = \mathrm{span}\{f_1,\ldots,f_K\}$, the least-squares estimator and test statistic
-\begin{equation*}
-\wt{f}_{\mathrm{LS}} := \argmin_{f \in F_K} \|Y - f\|_n^2,\textrm{and}~~\wt{T}_{\mathrm{LS}} = \|\wt{f}\|_{\nu}^2
-\end{equation*}
-are still rate-optimal over $H_0^s(P)$.
-However, this is not a totally satisfactory fix.
-For one thing, the least squares approach still requires that we know $\Delta$, and is not well suited for the case where $\mc{X}$ is a manifold, and $\Delta$ is the manifold Laplace-Beltrami operator.
-Additionally, diagonalizing even the unweighted Laplace-Beltrami operator $\Delta$ is quite difficult for all but a few special domains, such as $\mc{X} = [0,1]^d$.
-The general message is that in order to be minimax optimal, classical spectral series and least-squares estimators require an unrealistic amount of information on the design, and can be implemented only in selective situations.
-
-In contrast, Laplacian eigenmaps regression uses the eigenvectors of the neighborhood graph Laplacian $L_{n,\varepsilon}$ as features.
-The neighborhood graph Laplacian can be constructed (basically) without any knowledge of $P$ or $\mc{X}$, and Laplacian eigenmaps is thus a practicable method for regression.
-Moreover, the eigenvectors of $L_{n,\varepsilon}$ are approximations to the eigenfunctions $f_1,f_2,\ldots$ of $\Delta_P$---as we will discuss momentarily---so that $\wh{f}$ and $\wh{T}$ can in turn be viewed as approximations to the $\wt{f}$ and $\wt{T}$, respectively.
-As this view suggests, Laplacian eigenmaps naturally incurs some additional error compared to traditional spectral series estimators---this is the price we pay for using an approximation of $\Delta_P$.
-Fortunately, the general message of this paper is that despite this extra error, Laplacian eigenmaps retains many of the favorable optimality properties of its classical spectral series counterparts while freeing us from requiring unrealistic knowledge of $\Delta_P$.
-
-\subsection{Our contributions}
-The following items summarize our major results.
-In all of them we assume~\ref{asmp:domain} and~\ref{asmp:density} hold, but not that any of $P$, $\mc{X}$, $\Delta_P$ or $\Delta$ are known.
-Put $r := s - 1$.
-\begin{itemize}
-	\item \textbf{Minimax optimal estimation.} For any $f_0 \in H^1(\Xset;M)$, with high probability $\norm{\wh{f} - f_0}_{n}^2 \lesssim M^{2d/(2 + d)}n^{-2/(2 + d)}$.
-	\item \textbf{Minimax optimal testing.}
-	A test constructed using $\wh{T}$ has non-trivial power whenever $f_0 \in H^1(\Xset;M)$ satisfies $\norm{f_0}_{P}^2 \gtrsim M^{2d/(2 + d)}n^{-4/(4 + d)}$ and $d < 4$.
-	\item \textbf{Higher-order smoothness conditions.} When the regression function $f_0 \in H_0^s(\Xset;M)$ and additionally the density $p \in C^{r}(\mc{X};M)$, with high probability the error $\norm{\wh{f} - f_0}_{n}^2 \lesssim M^{2d/(2s + d)}n^{-2s/(2s + d)}$.
-Additionally, a test constructed using $\wh{T}$ has non-trivial power whenever $f_0 \in H_0^s(\Xset;M)$ satisfies $\norm{f_0}_{P}^2 \gtrsim M^{2d/(4s + d)}n^{-4s/(4s + d)}$ and $d < 4$.
-	\item \textbf{Manifold adaptivity.}
-	If $\mc{X} \subset \Rd$ is a manifold of dimension $m < d$, and if $H^s(\mc{X})$ for $s \leq 4$, then each of the aforementioned rates hold with $d$ replaced by $m$.
-\end{itemize}
-The above estimation results are stated with respect to in-sample mean squared error, meaning the loss is measured in the empirical squared-norm $\|\cdot\|_n^2$.\footnote{To state the obvious: this is very different than measuring the \emph{training error} $\|Y - \wh{f}\|_n^2$.} This is natural, as Laplacian eigenmaps is defined only at the observed design points $X_1,\ldots,X_n$.
-That being said, often one would like an upper bound on the prediction risk, or equivalently the out-of-sample loss $\|\cdot\|_P^2$.
-We therefore also introduce a simple, intuitive way of extending $\wh{f}$ to a function defined over all of $\mc{X}$, and show that this extension is minimax rate-optimal with respect to error measured in $L^2(P)$ norm.
-
-\subsection{Related Work}
+\begin{table}
+	\begin{center}
+		\begin{tabular}{p{.2\textwidth} | p{.14\textwidth} p{.12\textwidth} }
+			Smoothness order & Flat Euclidean (Model~\ref{def:model_flat_euclidean}) & Manifold (Model~\ref{def:model_manifold}) \\
+			\hline
+			$s \leq 3$ & $\bm{n^{-2s/(2s + d)}}$ & $\bm{n^{-2s/(2s + m)}}$ \\
+			$s > 3$  & $\bm{n^{-2s/(2s + d)}}$ & \textcolor{red}{(?)}
+		\end{tabular}
+	\end{center}
+	\caption{Summary of estimation rates over Sobolev balls.
+Bold font marks minimax optimal rates.
+In each case, rates hold for all $d \in \mathbb{N}$ (under Model~\ref{def:model_flat_euclidean}), and for all $d, m \in \mathbb{N}$ (under Model~\ref{def:model_manifold}).
+Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $M \gtrsim n^{-1/2}$.}
+	\label{tbl:estimation_rates}
+\end{table}
+
+\begin{table}
+	\begin{center}
+		\begin{tabular}{p{.175\textwidth} p{.175\textwidth} | p{.14\textwidth} p{.12\textwidth} }
+			Smoothness order & Dimension & Flat Euclidean (Model~\ref{def:model_flat_euclidean}) & Manifold (Model~\ref{def:model_manifold}) \\
+			\hline
+			\multirow{2}{*}{$s = 1$} & $\dim(\mc{X}) < 4$ & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
+			& $\dim(\mc{X}) \geq 4$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
+			\hline
+			\multirow{3}{*}{$s = 2$ or $3$} & $\dim(\mc{X}) \leq 4$  & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
+			& $4 <\dim(\mc{X}) < 4s$  & \textcolor{red}{(?)}& \textcolor{red}{(?)} \\
+			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
+			\hline
+			\multirow{3}{*}{$s > 3$} & $\dim(\mc{X}) \leq 4$ & $\bm{n^{-4s/(4s + d)}}$ & \textcolor{red}{(?)} \\
+			& $4 < \dim(\mc{X}) < 4s$ & \textcolor{red}{(?)} & \textcolor{red}{(?)} \\
+			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
+		\end{tabular}
+	\end{center}
+	\caption{Summary of testing rates over Sobolev balls.
+Bold font marks minimax optimal rates.
+Rates when $d > 4s$ assume that $f_0 \in L^4(P,M)$.
+Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $M \gtrsim n^{-1/2}$.}
+	\label{tbl:testing_rates}
+\end{table}
+
+\paragraph{Related Work.} 
 \begin{itemize}
 	\item \textcolor{red}{(TODO)} Summarize what is known on fixed graphs.
-	\item \textcolor{red}{Belkin, Slepcev} show that for a fixed $k$, the eigenvector $v_k$ converges to the eigenfunction $u_k$, in the sense that \textcolor{red}{(...)}.
+	\item \textcolor{red}{von Luxburg, Belkin, Slepcev} show that for a fixed index $k$, the eigenvector $v_k$ converges to the eigenfunction $u_k$, in the sense that \textcolor{red}{(...)}.
 \textcolor{red}{Shi, Burago, Garcia Trillos, Calder, Cheng} build on this, giving finite sample bounds, rates of convergence, and making statements uniform over $1 \leq k \leq k_{\max}$.
-This justifies our previous discussion, in which we treat Laplacian eigenmaps as an approximation of spectral series methods.
-However, we emphasize that the minimax optimality of $\wh{f}$ or a test based on $\wh{T}$ does not follow straightforwardly from such results.
-Instead our analysis proceeds along different lines, which we detail in  Section~\ref{sec:analysis}.
+This suggests that Laplacian eigenmaps may be good alternatives to classical spectral series methods when the latter cannot be implemented (e.g.
+because $P$ or $\mc{X}$ are unknown.)
 	\item There has been limited analysis of the estimator $\wh{f}$.
@@
 In this case the estimator reduces to a classical spectral series estimator, and there is no error incurred by approximating $\Delta_P$.
-\cite{lee2016} consider a similar estimator---using eigenvectors of a different normalization of the Laplacian $L$---in both the supervised and semi-supervised setups, but get suboptimal rates for the supervised problem.
+\cite{lee2016} consider the \emph{diffusion maps} estimator---which uses the eigenvectors of a different normalization of the Laplacian $L$---in both the supervised and semi-supervised setups, but they derive suboptimal rates for the supervised problem.
 As far as we know, there has been no analysis of the test statistic $\wh{T}$ in the random design framework which we study.
-	\item \textcolor{red}{(TODO)}: Alternatively, \cite{trillos2020} \textcolor{blue}{(Green 2021)} use the graph Laplacian operator to induce a penalty over functions $f: \{\bf X\} \to \Reals$, and study a regularized estimator estimator, which minimizes the penalized loss
-	\begin{equation*}
-	\|Y - f\|_n^2 + \lambda \dotp{L_{n,\varepsilon}f}{f}_n.
-	\end{equation*}
-	\textcolor{blue}{(Green 2021)} show that the resulting estimator is (nearly)-minimax optimal, but only for $s = 1$ and $d \leq 4$.
+	\item \textcolor{red}{(TODO)}: Alternatively, \cite{trillos2020} \textcolor{blue}{(Green 2021)} use the graph Laplacian operator to induce a penalty over functions $f: \{\bf X\} \to \Reals$.
+The \emph{Laplacian smoothing} estimator $\wh{f}_{\mathrm{LS}}$ is obtained by minimizing the sum of this penalty along with a data-fidelity term,
+	\begin{equation*}
+	\wh{f}_{\mathrm{LS}} := \|Y - f\|_n^2 + \lambda \dotp{L_{n,\varepsilon}f}{f}_n.
+	\end{equation*}
+	\textcolor{blue}{(Green 2021)} show that the resulting estimator is minimax optimal, but only for $s = 1$ and $d \leq 4$.
 In contrast, the Laplacian eigenmaps estimator is optimal for all $s$ and $d$.
@@
 
+\textbf{(5) Organization.}
+
+\textcolor{red}{(TODO)}
+
+\textbf{(6) Notation.}
+
+\textcolor{red}{TODO}:
+
+\begin{itemize}
+	\item Define $\Leb^2(\mc{X}), \Leb^2(P)$ and $\Leb^2(P_n)$. Define $C^k(\mc{X})$, its norm, and its normed ball.
+	\item Define asymptotic notation: $\ll$, $\lesssim$ and $\asymp$.
+\end{itemize}
+
+\section{Setup and Background}
+\label{sec:setup_main_results}
+
+In this section, we begin by giving a precise definition of our framework, and the Laplacian eigenmaps methods we study. We then review minimax rates for non-parametric regression over Sobolev spaces.
+
+\subsection{Non-parametric regression with random design}
+\label{sec:regression_laplacian_eigenmaps}
+
+We will operate in the usual setting of non-parametric regression with random design, in which we observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$.
+The design points $X_1,\ldots,X_n$ are sampled from a distribution $P$ with support $\mc{X} \subseteq \Rd$, and the responses follow
+\begin{equation}
+\label{eqn:model}
+Y_i = f_0(X_i) + w_i,
+\end{equation}
+with regression function $f_0: \mc{X} \to \Reals$, and $w_i \sim N(0,1)$ independent Gaussian noise.
+For simplicity we will assume throughout that the noise has unit-variance, but all of our results extend in a straightforward manner to the case where the variance is equal to a known positive value.
+
+We now formulate two models, which differ in the assumed nature of the support $\mc{X}$ of the design distribution $P$: the \emph{flat Euclidean} and \emph{manifold} models.
+
+\begin{definition}[Flat Euclidean model]
+	\label{def:model_flat_euclidean}
+	The data $(X_1,Y_1),\ldots,(X_n,Y_n)$ are sampled according to~\eqref{eqn:model}.
+The support $\mc{X}$ of the design distribution $P$ is an open, connected, and bounded subset of $\Rd$, with Lipschitz boundary.
+The distribution $P$ admits a density $p$ with respect to the $d$-dimensional Lebesgue measure $\nu$, which is bounded away from $0$ and $\infty$,
+	\begin{equation*}
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty, \quad \textrm{for all $x \in \mc{X}$.}
+	\end{equation*}
+\end{definition}
+
+In the following, recall that the injectivity radius of a $m$-dimensional Riemannian manifold $\mc{X}$ is the maximum value of $\delta$ such that the exponential map $\exp_x: B_m(0,\delta) \subset T_x(\mc{X}) \to B_{\mc{X}}(x,\delta) \subset \mc{X}$ is a diffeomorphism for all $x \in \mc{X}$.
+\begin{definition}[Manifold model]
+	\label{def:model_manifold}
+	The data $(X_1,Y_1),\ldots,(X_n,Y_n)$ are sampled according to~\eqref{eqn:model}. 
+	The support $\mc{X}$ of the design distribution $P$ is a closed, connected, smooth and boundaryless Riemannanian manifold embedded in $\Rd$, of intrinsic dimension $1 \leq m < d$.
+The injectivity radius of $\mc{X}$ is upper bounded by a constant $i_0$.
+The design distribution $P$ admits a density $p$ with respect to the volume form $d\mu$ induced by the Riemannian structure, which is bounded away from $0$ and $\infty$,
+	\begin{equation*}
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty, \quad \textrm{for all $x \in \mc{X}$.}
+	\end{equation*}
+\end{definition}
+
+\subsection{Laplacian eigenmaps}
+We now formally define the estimator and test statistic we study.
+Both are derived from eigenvectors of the neighborhood graph Laplacian operator.
+For a positive, symmetric kernel $\eta: [0,\infty) \to [0,\infty)$, and a bandwidth parameter $\varepsilon > 0$, the \emph{neighborhood graph Laplacian} operator $L_{n,\varepsilon}$ is
+\begin{equation}
+\label{eqn:neighborhood_graph_laplacian}
+L_{n,\varepsilon}u(x) = \frac{1}{n\varepsilon^{2 + \mathrm{dim}(\mc{X})}} \sum_{i = 1}^{n} \bigl(u(x) - u(X_j)\bigr) \eta\biggl(\frac{\|x - X_j\|}{\varepsilon}\biggr)
+\end{equation}
+(Here $\mathrm{dim}(\mc{X})$ stands for the dimension of $\mc{X}$.
+It is equal to $d$ under the assumptions of Model~\ref{def:model_flat_euclidean}, and equal to $m$ under the assumptions of Model~\ref{def:model_manifold}.
+The pre-factor $(n\varepsilon^{2 + \mathrm{dim}(\mc{X})})^{-1}$ is purely for convenience in taking limits as $n \to \infty, \varepsilon \to 0$).
+
+Although~\eqref{eqn:neighborhood_graph_laplacian} makes sense for any $x \in \mc{X}$ and $u: \mc{X} \to \Reals$, we will typically think of $x \in \{X_1,\ldots,X_n\}$ and $u = (u(X_1),\ldots,u(X_n)) \in \Leb^2(P_n)$), so that $L_{n,\varepsilon}$ is equivalently a matrix in $\Reals^{n \times n}$.
+Indeed, viewed as a matrix $L_{n,\varepsilon}$ is a symmetric and positive semi-definite, and can thus be diagonalized by its eigenpairs, solutions $(\lambda,v)$ to the eigenproblem
+\begin{equation*}
+\frac{1}{n}L_{n,\varepsilon}v = \lambda v, \quad \|v\|_n^2 = 1.
+\end{equation*}
+We will assume without loss of generality that each eigenvalue $\lambda$ of $L_{n,\varepsilon}$ has algebraic multiplicity $1$, and so we can index the eigenpairs $(\lambda_1,v_1),\ldots,(\lambda_n,v_n)$ in ascending order of eigenvalue, $0 = \lambda_1 < \ldots < \lambda_n$.
+
+The Laplacian eigenmaps estimator $\wh{f}$ simply projects the response vector ${\bf Y}$ onto the first $K$ eigenvectors of $L_{n,\varepsilon}$: letting $V_K \in \Reals^{n \times K}$ be the matrix with columns $v_1,\ldots,v_K$, we have that
+\begin{equation}
+\label{eqn:laplacian_eigenmaps_estimator}
+\wh{f} := \sum_{k = 1}^{K} \dotp{Y}{v_k}_{n} v_k = \frac{1}{n} V_K V_K^{\top} Y.
+\end{equation} 
+If $\wh{f}$ is a reasonable estimate of $f_0$, then the Laplacian eigenmaps test statistic
+\begin{equation}
+\label{eqn:laplacian_eigenmaps_test}
+\wh{T} := \|\wh{f}\|_n^2
+\end{equation}
+is in turn a reasonable estimate of $\|f_0\|_{P}^2$, and can be used to distinguish whether or not $f_0 = 0$.
+
+It may be helpful to comment briefly on the term ``Laplacian eigenmaps'', which we use a bit differently than is typical in the literature.
+Laplacian eigenmaps typically refers to an algorithm for embedding, which maps each design point $X_1,\ldots,X_n$ to $\Reals^K$ according to $X_i \mapsto (v_{1,i}, \ldots, v_{K,i})$.
+Viewing this mapping as a feature map, we can then interpret the estimator $\wh{f}$ as the least-squares solution to a linear regression problem with responses $Y_1,\ldots,Y_n$ and features $v_1,\ldots,v_K$.
+Often, the Laplacian eigenmaps embedding is viewed as a tool for dimensionality reduction, wherein it is implicitly assumed that $K \ll d$.
+We will neither explicitly nor implicitly take $K < d$; indeed, the embedding perspective is not particularly illuminating in what follows, and we do not henceforth make reference to it.
+Instead, we use ``Laplacian eigenmaps'' to directly refer to the estimator $\wh{f}$ or test statistic $\wh{T}$.
+
+On a different note, it is worth explicitly pointing out that $\wh{f}$ is defined only in-sample, that is, only at the design points $X_1,\ldots,X_n$.
+There is no straightforward or agreed upon way to extend $\wh{f}$ out-of-sample, and we address the problem of out-of-sample estimation in some detail later in Section~\ref{sec:out_of_sample}.
+
+\subsection{Sobolev Classes}
+\label{sec:sobolev}
+Before we give an overview of our main results, we will first pause to review the definition of Sobolev classes.
+We divide our discussion into two cases---the flat Euclidean case, and the manifold case, corresponding to Models~\ref{def:model_flat_euclidean} and Model~\ref{def:model_manifold}.
+
+\paragraph{Flat Euclidean case.}
+Take $\mc{X}$ to be an open, connected set with Lipschitz boundary, as in Model~\ref{def:model_flat_euclidean}.
+Let $Q$ be a measure supported on a subset of $\mc{X}$, that is absolutely continuous with respect to the full-dimensional Lebesgue measure $\nu$.
+Recall that for a given multiindex $\alpha \in \mathbb{N}^d$, a function $f$ is \emph{$\alpha$ weakly differentiable} if there exists some $h \in L^1(Q)$ such that
+\begin{equation*}
+\int_{\mc{X}} h g = (-1)^{|\alpha|} \int_{\mc{X}} f D^{\alpha}g, \quad \textrm{for every $g \in C_0^{\infty}(\mc{X})$.}
+\end{equation*}
+If such a function $h$ exists, it is the $\alpha$th weak partial derivative of $f$, and denoted by $D^{\alpha}f = h$.
+If $f$ is $\alpha$ weakly differentiable for all $|\alpha| = j$, then $d^jf := \sum_{|\alpha| = j}D^{\alpha}f$ is the sum of all weak partial derivatives of order $j$.
+
+For an integer $s \geq 1$, the Sobolev class $H^s(Q)$ consists of all those functions $f \in \Leb^2(Q)$ which have weak derivatives of all orders $j = 1,\ldots,s$, all of which also belong to $\Leb^2(Q)$.
+The $j$th order semi-norm for $f \in H^s(Q)$ is $|f|_{H^j(Q)} := \|d^jf\|_{L^2(Q)}$.
+The corresponding norm
+\begin{equation*}
+\|f\|_{H^s(Q)}^2 := \|f\|_{\Leb^2(Q)}^2 + \sum_{j = 1}^{s} |f|_{H^j(Q)}^2,
+\end{equation*}
+which induces the ball
+\begin{equation*}
+H^s(Q;M) := \bigl\{f \in H^s(Q): \|f\|_{H^s(Q)} \leq M\bigr\}.
+\end{equation*} 
+In the special case where $Q$ is the uniform measure over $\mc{X}$, we use the notation $H^s(\mc{X}) := H^s(Q)$.
+Finally, we note that $H^s(Q)$ is the completion of $C^{(\infty)}(\mc{X})$ with respect to the $\|\cdot\|_{H^s(Q)}$ norm, so that $C^{\infty}(\mc{X})$ is dense in $H^s(\mc{X})$.
+
+\paragraph{Manifold case.}
+
+\textcolor{red}{(TODO)}
+
+\paragraph{Boundary conditions.}
+In the flat Euclidean case, when $s > 1$ we will assume that $f_0$ satisfies some boundary conditions.
+In particular, we will assume that $f_0$ belongs to the set of \emph{zero trace} Sobolev functions $H_0^{s}(P)$, defined as the closure of $C_0^s(\mc{X})$ with respect to $\|\cdot\|_{H^s(P)}$ norm.
+Of course, under Model~\ref{def:model_manifold} the domain $\mc{X}$ is without boundary, and so boundary conditions are irrelevant.
+
+Let us explain the reason we require boundary conditions in the flat Euclidean case when $s > 1$, by considering an alternative way to characterize Sobolev spaces using a relation to sequence spaces.\footnote{This is equivalent to characterizing Sobolev spaces using the Fourier transform, but we find it more transparent to instead work directly with sequence spaces.} Let $(\lambda_1(\Delta_P),\psi_1),(\lambda_2(\Delta_P),\psi_2),\ldots$ be the solutions to the Laplace-Beltrami eigenvector equation (with Neumann boundary conditions),
+\begin{equation*}
+\Delta_P\psi_k = \lambda_{k}(\Delta_P) \psi_k, \quad \frac{\partial}{\partial {\bf n}}\psi_k = 0~~\textrm{on $\partial \mc{X}$},
+\end{equation*}
+where $\partial/(\partial {\bf n})$ is the partial derivative operator in the direction of the normal vector $\mathbf{n}$ to the boundary of $\mc{X}$.
+We define an ellipsoid $\Theta \subset \ell^2(\mathbb{N})$  using the eigenvalues $\lambda_1(\Delta_P),\lambda_2(\Delta_P),\ldots$ as $\Theta := \{\theta \in \ell^2(\mathbb{N}): \sum_{k = 1}^{\infty} \lambda_k(\Delta_P)^{2s} \theta_k^2 < \infty\}$.
+Then, the function class $\mc{H}^s(P) = \{f = \sum_{k = 1}^{\infty} \theta_k \psi_k: \theta \in \Theta\}$ consists precisely of those Sobolev functions $f \in H^s(P)$ which satisfy the following \emph{Neumann} boundary conditions,
+\begin{equation}
+\label{eqn:condition}
+\frac{\partial^j}{\partial {\bf n}^j}f = 0 \quad \textrm{for $j = 1,3,5,\ldots,\floor{s/2}$.}
+\end{equation}
+As we have already reviewed, the eigenvectors $v_k$ of the graph Laplacian $L_{n,\varepsilon}$ converge to $\psi_k$ as $n \to \infty, \varepsilon \to 0$.
+As a result, is natural that they accurately approximate only those functions $f_0 \in \mc{H}^s(P) \subset H^s(P)$.
+Although the zero-trace boundary condition is more restrictive than~\eqref{eqn:condition}---since it also requires that the even derivatives be equal to $0$---the point is that some kind of boundary condition will be necessary in order to obtain optimal rates for spectral series methods, including Laplacian eigenmaps.
+We will stick to the zero-trace condition, which greatly eases some steps in our proofs.
+
+\subsection{Minimax Rates}
+\label{subsec:minimax_rates_sobolev}
+We now review the minimax estimation and goodness-of-fit testing rates over Sobolev balls.
+We pay special attention to the spectral series methods which achieve these rates, because, as we have already seen, these methods are very related to Laplacian eigenmaps.
+
+\paragraph{Estimation rates over Sobolev balls.}
+In the estimation problem, we ask for an estimator---formally, a Borel measurable function $\wh{f}$, that maps from $\mc{X}$ to $\Reals$---which is close to the regression function $f_0$ with respect to the squared norm $\|\wh{f} - f_0\|_{P}^2$.
+Under Model~\ref{def:model_flat_euclidean}, the minimax estimation rate over the Sobolev ball is\footnote{Perhaps surprisingly, we were unable to find a reference that explicitly states the rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate} in our general setup, for all values of $s$ and $d$---but the rate certainly does hold.
+Momentarily we will introduce a spectral series estimator $\wt{f}$ which achieves this rate, implying the upper bound.
+The lower bound in~\eqref{eqn:sobolev_space_minimax_estimation_rate} can be obtained from the analogous lower bound over $C_0^s(\mc{X})$\textcolor{red}{(reference)}, and the fact that there exists a continuous embedding of $C_0^s(\mc{X})$ into $\mc{H}^s(P)$.}
+\begin{equation}
+\label{eqn:sobolev_space_minimax_estimation_rate}
+\inf_{\wh{f}} \sup_{f_0} \|\wh{f} - f_0\|_P^2 \asymp M^2(M^2n)^{-2s/(2s + d)}
+\end{equation}
+here the infimum is taken over all estimators $\wh{f}$, the supremum over all $f_0 \in H^1(\mc{X};M)$ (first-order case) or $f_0 \in H_0^s(\mc{X};M)$ (higher-order case).
+
+The upper bound in~\eqref{eqn:sobolev_space_minimax_estimation_rate} can be certified by a particular spectral series estimator.
+Under the assumptions of Model~\ref{def:model_flat_euclidean}, the eigenvalues $\lambda_k(\Delta_P)$ satisfy the Weyl's Law asymptotic scaling $\lambda_{k}(\Delta_P) \asymp k^{2/d}$, a fact that we discuss in more detail later.
+A key consequence of this scaling is that the Sobolev ball $H_0^s(P) \subset \mc{H}^s(P)$ continuously embeds into the Sobolev ellipsoid $\mc{\wt{H}}^s(P) = \{f = \sum_{k = 1}^{\infty} \theta_k \psi_k: \sum_{k = 1}^{\infty} (k - 1)^{2s/d} \theta_k^2\}$.
+By reducing the regression problem to an estimation problem in a Gaussian sequence model \textcolor{red}{(see e.g.
+Tsybakov,Johnstone,Gine)}, it can be shown the estimator
+\begin{equation*}
+\wt{f} := \sum_{k = 1}^{K} \dotp{Y}{\psi_k}_n \psi_k,
+\end{equation*}
+achieves the minimax rate over $\mc{H}^s(P;M)$, and the same upper bound on the minimax rate therefore applies to $H_0^s(P;M)$.
+
+\paragraph{Goodness-of-fit testing rates over Sobolev balls.}
+In the goodness-of-fit testing problem, we ask for a test function---formally, a Borel measurable function $\phi$ that takes values in $\{0,1\}$--- which can distinguish between the hypotheses
+\begin{equation}
+\mathbf{H}_0: f_0 = f_0^{\star}, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{f_0^{\star}\}.
+\end{equation} 
+Typically, the null hypothesis $f_0 = f_0^{\star} \in \mc{F}$ reflects the absence of interesting structure, and $\mc{F} \setminus  \{f_0^{\star}\}$ is a set of smooth departures from this null.
+To fix ideas, as in \citet{ingster2009} we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\star} = 0$ and $\mc{F} = \mc{H}^s(\mc{X};M)$ is a Sobolev ball.
+This is without loss of generality since our test statistic and its analysis are easily modified to handle the case when $f_0^{\star}$ is not $0$, by simply subtracting $f_0^{\star}(X_i)$ from each observation $Y_i$.
+
+The Type I error of a test $\phi$ is $\mathbb{E}_0[\phi]$, and if $\mathbb{E}_0[\phi] \leq a$ for a given $a \in (0,1)$ we refer to $\phi$ as a level-$a$ test\footnote{We reserve the more common symbol $\alpha$ for multi-indices, so as to avoid confusion}.
+The worst-case risk of $\phi$ over $\mc{F}$ is
+\begin{equation*}
+R_n(\phi, \mc{F}, \epsilon) := \sup\Bigl\{\mathbb{E}_{f_0}[1 - \phi]: f_0 \in \mc{F}, \|f_0\|_{P} > \epsilon\Bigr\},
+\end{equation*}
+and for a given constant $b \in (0,1)$, the minimax critical radius $\epsilon(\mc{F})$ is the smallest value of $\epsilon$ such that some level-$a$ test has worst-case risk of at most $b$. Formally,
+\begin{equation*}
+\epsilon(\mc{F}) := \inf\Bigl\{\epsilon > 0: \inf_{\phi} R_n(\phi,\mc{F},\epsilon) \leq b \Bigr\},
+\end{equation*} 
+where in the above the infimum is over all level-$a$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Clearly, the minimax critical radius $\epsilon(\mc{F})$ depends on $a$ and $b$.
+However, we adopt the typical convention of treating $\alpha,b \in (0,1)$ and  as small but fixed positive constants; hence they will not affect the testing error rates, and we suppress them notationally.} See \citet{ingster82,ingster87,ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
+
+Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing critical radius is much smaller than the minimax estimation rate.
+\citet{ingster2009} show that in the special case where $\mc{X} = [0,1]^d$ and $P = \nu$,
+\begin{equation}
+\label{eqn:sobolev_space_testing_critical_radius}
+\epsilon^2\bigl(\mc{H}^s(P;M)\bigr) \asymp M^2(M^2n)^{-4s/(4s + d)}~~\textrm{for $1 \leq d < 4s$,}
+\end{equation}
+and the same result holds under the more general setup of Model~\ref{def:model_flat_euclidean}.
+The test \citet{ingster2009} use to certify the upper bound in~\eqref{eqn:sobolev_space_testing_critical_radius} is a spectral series method, which uses the squared $L^2(P)$ norm of $\wt{T} = \|\wt{f}\|_{P}^2$ as a test statistic.
+
+When $4s \geq d$ the functions in $H^s(\Xset)$ are very irregular---formally speaking $H^s(\Xset)$ does not continuously embed into $\Leb^4(\Xset)$ when $4s \geq d$---and as far as we know, the minimax testing rates in this regime have not been worked out.
+However, if one explicitly assumes that $f_0 \in \Leb^4(\mc{X})$ then the critical radius is given by $\epsilon^2 \asymp Mn^{-1/2}$; as a sanity check, note that this is strictly worse than the rate in~\eqref{eqn:sobolev_space_testing_critical_radius}.
+As we discuss after our first main theorem regarding testing (Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}), this rate is achievable by a test based on $\wh{T}$.
+
+\paragraph{Manifold setup.}
+Under Model~\ref{def:model_manifold}, both~\eqref{eqn:sobolev_space_minimax_estimation_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced everywhere by the intrinsic dimension $m$.
+The estimator $\wt{f}$ and a test using the statistic $\wt{T}$---with $\psi_1,\psi_2,\ldots$ now the eigenfunctions of the manifold weighted Laplace-Beltrami operator $\Delta_P$---continue to achieve the optimal rates.
+\textcolor{red}{(TODO): Find references.}
+
+\paragraph{In-sample mean squared error.}
+
+The error in~\eqref{eqn:sobolev_space_minimax_estimation_rate} is measured with respect to $\|\cdot\|_P^2$ norm.
+Since $\wh{f}$ is defined only at the design points $X_1,\ldots,X_n$, we will measure estimation rates in $\|\cdot\|_n^2$ error.
+The lower bounds used to show~\eqref{eqn:sobolev_space_minimax_estimation_rate} will still hold when in-sample loss is used rather than out-of-sample loss \textcolor{red}{(?)}, and we verify matching upper bounds in the sections to come.
+Thus the minimax rates are unchanged---we call estimators which achieve these optimal rates \emph{in-sample minimax rate-optimal estimators}.
+
+% TODO: AG, get rid of this!
+
+%\subsection{Our contributions}
+%The following items summarize our major results.
+In all of them we assume~\ref{asmp:domain} and~\ref{asmp:density} hold, but not that any of $P$, $\mc{X}$, $\Delta_P$ or $\Delta$ are known.
+Put $r := s - 1$.
+%\begin{itemize}
+%	\item \textbf{Minimax optimal estimation.} For any $f_0 \in H^1(\Xset;M)$, with high probability $\norm{\wh{f} - f_0}_{n}^2 \lesssim M^{2d/(2 + d)}n^{-2/(2 + d)}$.
+%	\item \textbf{Minimax optimal testing.}
+%	A test constructed using $\wh{T}$ has non-trivial power whenever $f_0 \in H^1(\Xset;M)$ satisfies $\norm{f_0}_{P}^2 \gtrsim M^{2d/(2 + d)}n^{-4/(4 + d)}$ and $d < 4$.
+%	\item \textbf{Higher-order smoothness conditions.} When the regression function $f_0 \in H_0^s(\Xset;M)$ and additionally the density $p \in C^{r}(\mc{X};M)$, with high probability the error $\norm{\wh{f} - f_0}_{n}^2 \lesssim M^{2d/(2s + d)}n^{-2s/(2s + d)}$.
+Additionally, a test constructed using $\wh{T}$ has non-trivial power whenever $f_0 \in H_0^s(\Xset;M)$ satisfies $\norm{f_0}_{P}^2 \gtrsim M^{2d/(4s + d)}n^{-4s/(4s + d)}$ and $d < 4$.
+%	\item \textbf{Manifold adaptivity.}
+%	If $\mc{X} \subset \Rd$ is a manifold of dimension $m < d$, and if $H^s(\mc{X})$ for $s \leq 4$, then each of the aforementioned rates hold with $d$ replaced by $m$.
+%\end{itemize}
+%In the statements above, estimation error is measured in-sample, using the empirical squared-norm $\|\cdot\|_n^2$.\footnote{To state the obvious: this is very different than measuring the \emph{training error} $\|Y - \wh{f}\|_n^2$.} It is necessary to measure loss in-sample because the Laplacian eigenmaps estimator $\wh{f}$ is defined only at the observed design points $X_1,\ldots,X_n$.
+That being said, often one would like an upper bound on the prediction risk, or equivalently the out-of-sample loss $\|\cdot\|_P^2$.
+We therefore also introduce a simple, intuitive way of extending $\wh{f}$ to a function defined over all of $\mc{X}$, and show that this extension is minimax rate-optimal with respect to error measured in $L^2(P)$ norm.
+
 \section{Minimax Optimality of Laplacian Eigenmaps}
@@
 
-In this section, we will show that the estimator $\wh{f}$, and a test using the statistic $\wh{T}$, achieve optimal estimation and goodness-of-fit testing rates over Sobolev classes.
-We will divide our statements based on whether $f_0$ belongs to the first order Sobolev class ($s = 1$) or a higher-order Sobolev class ($s > 1$), since the details of the two settings are somewhat different.
-Throughout this section, we will always assume~\ref{asmp:domain} and~\ref{asmp:density}.
+As previously explained, Laplacian eigenmaps is a discrete and noisy approximation to a spectral series method, of particular use when the design distribution $P$ is unknown.
+The hope is that the error incurred by this approximation is sufficiently small such that Laplacian eigenmaps preserves the optimality properties of the spectral series methods they are approximating.
+In this section and the following one, we will show that this is indeed the case; the estimator $\wh{f}$, and a test using the statistic $\wh{T}$, achieve optimal estimation and goodness-of-fit testing rates over Sobolev classes.
+
+In this Section we will cover the flat Euclidean case, where we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to Model~\ref{def:model_flat_euclidean}.
+We will divide our theorem statements based on whether we assume the regression function $f_0$ belongs to the first order Sobolev class ($s = 1$) or a higher-order Sobolev class ($s > 1$), since the details of the two settings are somewhat different.
 
@@
 \label{sec:first_order_sobolev_classes}
-To begin, we handle the case where $f_0 \in H^1(P)$.
+We begin by assuming $f_0 \in H^1(P)$.
 We show that $\wh{f}$ and a test based on $\wh{T}$ are minimax optimal, for all values of $d$, and under no additional assumptions on the data generating process, i.e.
@@
 
-\paragraph{Estimation.} When the graph radius $\varepsilon$ and number of eigenvectors $K$ are chosen appropriately, the estimator $\wh{f}$ has (in-sample) optimal risk (up to constant factors) $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2/(2 + d)}$.
+\paragraph{Estimation.} When the graph radius $\varepsilon$ and number of eigenvectors $K$ are chosen appropriately, we show in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that the estimator $\wh{f}$ achieves the minimax rate, when error is measured in $\|\cdot\|_{n}$.
 \begin{enumerate}[label=(A\arabic*)]
@@
 	\label{eqn:radius_fo} 
-	C_0 \biggl(\frac{\log n}{n}\biggr)^{1/d} \leq \varepsilon \leq \min\{i_0,K^{-1/d}\},
+	\biggl(\frac{\log n}{n}\biggr)^{1/d} \ll \varepsilon \ll \min\{1,K^{-1/d}\},
 	\end{equation}
@@
 	\label{eqn:eigenvector_estimation_fo} 
-	c_1 (M^2 n)^{2d/(2 + d)}\leq K \leq C_1 (M^2 n)^{2d/(2 + d)}.
+	K \asymp \min\Bigl\{(M^2n)^{d/(2 + d)} \vee 1, n\Bigr\}
 	\end{equation}
@@
 	\label{thm:laplacian_eigenmaps_estimation_fo}
-	Suppose $f_0 \in H^1(P,M)$.
-If the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_fo}, then the following statement holds for any $\delta \in (0,1)$: with probability at least $1 - \delta - \textcolor{red}{(???)}$,
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H^1(P,M)$.
+There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_fo}, then
 	\begin{equation}
 	\label{eqn:laplacian_eigenmaps_estimation_fo}
-	\|\wh{f} - f_0\|_n^2 \leq \frac{C}{\delta}M^{2d/(2 + d)}n^{-2/(2 + d)}
-	\end{equation}
+	\|\wh{f} - f_0\|_n^2 \leq C\Bigl(\frac{1}{\delta}M^2(M^2n)^{-2/(2 + d)} \wedge 1\Bigr) \vee \frac{1}{n},
+	\end{equation}
+	with probability at least $1 - \delta - Cn\exp(-cn\varepsilon^d)$.
 \end{theorem}
-Some remarks:
+From~\eqref{eqn:laplacian_eigenmaps_estimation_fo} it follows immediately that with high probability $\|\wh{f} - f_0\|_n^2 \lesssim M^2(M^2n)^{-2/(2 + d)}$ whenever $n^{-1/2} \lesssim M \lesssim n^{1/d}$, and is thus an in-sample minimax rate-optimal estimator.
+
+Some other remarks:
 \begin{itemize}
-	\item Regarding the restrictions on hyperparameters: the lower bound on $\varepsilon$ is the connectivity threshold, the smallest length scale at which the resulting graph will still, with high probability, be connected.
-On the other hand, as we will see in Section~\ref{subsec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K \asymp \lambda_K(\Delta_P) \asymp K^{2/d}$.
-Finally, the restriction $K \asymp (M^2n)^{2d/(2 + d)}$ is chosen  to optimally trade-off bias and variance.
-Together, these restrictions suggest that the
-	\item \textbf{(1.1) How to tune hyper-parameters.} We note that the ranges~\eqref{eqn:radius_fo} and~\eqref{eqn:eigenvector_estimation_fo} depend on unknown quantities such as the dimension $d$ and radius of the Sobolev ball $M$.
+	\item When $M \ll n^{-1/2}$ or $M \gg n^{1/d}$, the rate in~\eqref{eqn:laplacian_eigenmaps_estimation_fo} looks different than~\eqref{eqn:sobolev_space_minimax_estimation_rate}.
+In particular, when $M \ll n^{-1/2}$, then computing Laplacian eigenmaps with $K = 1$ achieves the parametric rate $\|\wh{f} - f_0\|_n^2 \lesssim n^{-1}$, and the zero-estimator $\wh{f} = 0$ achieves the better rate $\|\wh{f} - f_0\|_n^2 \lesssim M^2$, but neither match the minimax rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}.
+On the other hand, when $M \gg n^{1/d}$, then computing Laplacian eigenmaps with $K = n$ achieves the rate $\|\wh{f} - f_0\|_n^2 \asymp 1$, which is better than the rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}.
+However, in truth these are edge cases, \textcolor{red}{(...)}.
+	\item The lower bound on $\varepsilon$ imposed by~\eqref{eqn:radius_fo} is on the order of the connectivity threshold, the smallest length scale at which the resulting graph will still, with high probability, be connected.
+On the other hand, as we will see in Section~\ref{subsec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K$ is on the same order as the continuum eigenvalue $\lambda_K(\Delta_P)$.
+Finally, we choose $K \asymp (M^2n)^{2d/(2 + d)}$ to optimally trade-off bias and variance.
+	\item We note that the ranges~\eqref{eqn:radius_fo} and~\eqref{eqn:eigenvector_estimation_fo} depend on unknown quantities such as the dimension $d$ and radius of the Sobolev ball $M$.
 In practice, one typically tunes hyper-parameters by sample-splitting or cross-validation.
-However, because the estimator $\wh{f}$ is defined only in-sample, such approaches cannot be straightforwardly applied to select the graph radius $\varepsilon$, or number of eigenvectors $K$.
+However, because the estimator $\wh{f}$ is defined only in-sample, we cannot use such methods to select the graph radius $\varepsilon$, or number of eigenvectors $K$.
 We return to this issue upon considering out-of-sample extensions of $\wh{f}$ later, in Section~\ref{sec:out_of_sample}.
-	\item \textbf{(1.2) The dependence on failure probability can be improved.} The upper bound on in-sample risk given by Equation~\eqref{eqn:laplacian_eigenmaps_estimation_fo} depends on the pre-factor $C/\delta$, and holds with probability $1 - \delta - \textcolor{red}{(?)}$.
-When $f_0 \in C^1(\mc{X};M)$, we can improve these dependencies, changing the pre-factor in~\eqref{eqn:laplacian_eigenmaps_estimation_fo} to $C(1 + 1/\delta)$, and showing that the bound holds with probability at least $1 - C\delta^2/n$.
+	\item The upper bound given in Equation~\eqref{eqn:laplacian_eigenmaps_estimation_fo} holds with probability $1 - \delta - Cn\exp(-cn\varepsilon^d)$.
+Under the stronger assumption that $f_0 \in C^1(\mc{X};M)$ we can replace the factor of $\delta$ by the sharper $\delta^2/n$, which is less than $\delta$ because $\delta \in (0,1)$.
 \end{itemize}
 
-\textbf{(2) Laplacian eigenmaps is an optimal test.} Consider the test $\varphi = \1\{\wh{T} \geq t_{\delta}\}$, where $t_{\delta}$ is the threshold
-\begin{equation*}
-t_{\delta} := \frac{K}{n} + \frac{2K}{\sqrt{n}}.
-\end{equation*}
-The choice of $t_{\delta}$ guarantees that $\varphi$ is a level-$\delta$ test.
-Under appropriate choices of $\varepsilon$ and $K$, the test $\varphi$ has non-negligible power (up to constant factors) against alternatives separated the null by at least $\|f_0\|_{P}^2 \gtrsim M^{2d/(4 + d)}n^{-4/(4 + d)}$, whenever $d < 4$.
+\paragraph{Testing.} Consider the test $\varphi = \1\{\wh{T} \geq t_{a}\}$, where $t_{a}$ is the threshold
+\begin{equation*}
+t_{a} := \frac{K}{n} + \frac{1}{n}\sqrt{\frac{2K}{a}}
+\end{equation*}
+This choice of threshold $t_{a}$ guarantees that $\varphi$ is a level-$a$ test.
+Moreover, when $\varepsilon$ and $K$ are chosen appropriately, the test $\varphi$ has negligible Type II error against alternatives separated from the null by at least $\|f_0\|_{P}^2 \gtrsim M^2(M^2n)^{-4/(4 + d)}$, whenever $d < 4$.
 
@@
 	\label{eqn:eigenvector_testing_fo}
-	c_2 (M^2 n)^{2d/(4 + d)}\leq K \leq C_2 (M^2 n)^{2d/(4 + d)}.
+	K \asymp \min\Bigl\{(M^2n)^{2d/(4 + d)} \vee 1, n\Bigr\}
 	\end{equation}
 \end{enumerate}
-Let $p(d) = \min\{1,2/d\}$.
+In the following, let $p(d) = \min\{1,2/d\}$.
 \begin{theorem}
 	\label{thm:laplacian_eigenmaps_testing_fo}
-	Suppose $f_0 \in H^1(\mc{X};M)$.
-If the Laplacian eigenmaps test $\varphi$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_testing_fo}, then the following statement holds for any $\delta \in (0,1)$: if $f_0$ satisfies
+	Fix $a,b \in (0,1)$.
+Suppose Model~\ref{def:model_flat_euclidean}.
+Then $\mathbb{E}_0[\varphi] \leq a$, i.e $\varphi$ is a level-$a$ test.
+Suppose additionally $f_0 \in H^1(P,M)$, and that $d < 4$.
+Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if $f_0$ satisfies
 	\begin{equation}
 	\label{eqn:laplacian_eigenmaps_testing_criticalradius_fo}
-	\|f_0\|_P^2 \geq \frac{C}{\delta} \max\biggl\{\frac{1}{\delta}M^{2d/(4 + d)}n^{-4/(4 + d)}, M n^{-p(d)}\biggr\},
-	\end{equation}
-	then $\varphi$ has Type II error upper bounded by 
-	\begin{equation}
-	\Pbb_{f_0}(\varphi = 0) \leq \delta + \textcolor{red}{(...)}
-	\end{equation}
+	\|f_0\|_n^2 \geq \frac{C}{\delta}\biggl(\Bigl(M^2(M^2n)^{-4/(4 + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{1}{\delta n^{p(d)}} \biggr) \vee \frac{1}{n}
+	\end{equation}
+	then $\Ebb_{f_0}[1 - \phi] \leq b$.
 \end{theorem}
-Some remarks:
+Although~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} involves taking the maximum of several different terms, the most important takeaway is that if $n^{-1/2} \lesssim M^2 \ll n^{1/4}$ ($d = 1$) or $n^{-1/2} \lesssim M^2 \ll n^{2/d - 1/2}$ ($d = 2$ or $d = 3$), then $\varphi$ is a minimax rate-optimal test over $H^1(\mc{X},M)$.
+
+Some other remarks:
 \begin{itemize}
-	\item \textbf{(2.1) When $d \geq 4$, the testing problem is different.} When $d \geq 4$, it is not the case that $H^1(\mc{X})$ continuously embeds into $\Leb^4(\mc{X})$, and as far we know the optimal testing rates over $H^1(\mc{X})$ are not known.
-On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then simply taking $T = \|Y\|_n^2$ achieves the optimal rates.
-	\item \textbf{(2.2) Both of the remarks after Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} also apply to testing.}
+	\item When $d \geq 4$, as we have already mentioned $H^1(\mc{X})$ does not continuously embed into $\Leb^4(\mc{X})$.
+This has severe consequences for test statistics which involve taking an $L^2(P_n)$-norm, such as $\wh{T}$, or $L^2(P)$ norm, such as $\wt{T}$, since they may no longer have a finite variance.
+Indeed, as far we know when $d \geq 4$ the optimal testing rates over $H^1(\mc{X})$ are not known.
+On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then we show in Appendix~\textcolor{red}{(?)} that the Laplacian eigenmaps test, with $K = n$, is minimax optimal whenever $\|f_0\|_n^2 \gtrsim n^{-1/2}$.
+	\item \textcolor{red}{(TODO)}: Explain the presence of the $n^{p(d)}$ term. Explain that if you assume $f_0 \in C^1(\mc{X};M)$, you can avoid this term.
+	\item \textcolor{red}{(TODO)}: Explain what else you can get when you assume $C^1(\mc{X};M)$.
 \end{itemize}
@@
 \label{sec:higher_order_sobolev_classes}
-In this section, we now assume that the regression function displays some higher-order regularity, $f_0 \in H_0^s(\mc{X})$.
+In this section, we now assume that the regression function displays some higher-order regularity and is zero-trace, i.e.
+$f_0 \in H_0^s(P)$.
 We point out that on a graph there is no natural analogue to higher-order derivatives.
@@
 In the ``sub-critical'' regime $2s < d$, the Sobolev space $H^s(\mc{X})$ is quite irregular.
-For instance, it is not a Reproducing Kernel Hilbert Space, nor does it embed into any H\"{o}lder space.
+For instance, it is not a Reproducing Kernel Hilbert Space (RKHS), nor does it continuously embed into $C^0(\mc{X})$, much less into any H\"{o}lder space.
 As a result, certain versions of the nonparametric regression problem are ill-posed---for instance when loss is measured in $\Leb^{\infty}$ norm, or when the design points $\{X_1,\ldots,X_n\}$ are assumed to be fixed.
@@
 	\label{eqn:graph_seminorm_2}
-	\dotp{L^s f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^s} \|p\|_{C^{\beta}}
-	\end{equation}
-	When $s = 1$, the results hold for $f \in H^1(\mc{X})$ and any density $p$ that satisfies Assumption~\ref{asmp:density}.
+	\dotp{L^s f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^s} \|p\|_{C^{\beta}},
+	\end{equation}
+	with the same probability guarantee. When $s = 1$, the results hold for $f \in H^1(\mc{X})$ and any density $p$ that satisfies Assumption~\ref{asmp:density}.
 \end{proposition}
@@
 On the other hand, in order to give an estimate of $E_{\varepsilon}^{(s)}(f)$ in terms of the Sobolev norm $D_s(f)$, we first show that $L_{P,\varepsilon}^jf(x) \approx \Delta_P^jf(x)$, for $j = s/2$ when $s$ is odd and $j = (s - 1)/2$ when $s$ is even.
-In words, we show that a $j$th-order iterated difference operator approximates a $2j$-th (or $2j + 1$-th) order differential operator.
-In to establish this fact, we require that on the one hand $p \in C^{r}(\mc{X};M)$, and on the other hand $f$ be zero-trace.
+In words, we show that a $j$th-order iterated difference operator approximates a $2j$-th (or $2j + 1$-st) order differential operator.
+In to establish this fact, we require that on the one hand that $p \in C^{r}(\mc{X};M)$, and on the other hand that $f$ be zero-trace.
 
@@
 	\begin{equation*}
-	\lambda_k \geq c \cdot \min\Bigl\{\lambda_k(\Delta_P), \frac{1}{r^{2}} \Bigr\}
+	\lambda_k \geq c \cdot \min\Bigl\{\lambda_k(\Delta_P), \frac{1}{\varepsilon^{2}} \Bigr\}
 	\end{equation*}
 \end{proposition}
-By Weyl's Law, there exist constants $c_1$ and $C_1$ which do not depend on $k$, such that $c_1 k^{2/d} \leq \lambda_{k}(\Delta_P) \leq C_1 k^{2/d}$ for all $k \in \mathbb{N}$.
-As a result, $\lambda_{K} \geq c \lambda_{K}(\Delta_P)$ as long as $r \geq \kappa^{-1/d}$.
+By Weyl's Law \textcolor{red}{(Dunlop, Wu)}, there exist constants $c_1$ and $C_1$ which do not depend on $k$, such that $c_1 k^{2/d} \leq \lambda_{k}(\Delta_P) \leq C_1 k^{2/d}$ for all $k \in \mathbb{N}$.
+As a result, $\lambda_{K} \geq c \lambda_{K}(\Delta_P)$ as long as $\varepsilon \geq K^{-1/d}$.
 
@@
 
-It is also key that we directly analyze the semi-norm $\dotp{L^s f}{f}_n$, rather than invoking the pointwise convergence of $L^{s}f \to \Delta_P^{s}f$ to obtain a bound on $\dotp{L^s f}{f}_n$.
+It is also key that we directly analyze the semi-norm $\dotp{L_{n,\varepsilon}^s f}{f}_n$, rather than invoking the pointwise convergence of $L_{n,\varepsilon}^{s}f \to \Delta_P^{s}f$ in order to obtain a bound on $\dotp{L_{n,\varepsilon}^s f}{f}_n$.
 Recall that we assume $f$ has bounded derivatives up to (and including) order $s$.
@@
 The optimal rates for nonparametric regression suffer from a bona fide curse of dimensionality, and are thus difficult to reconcile with the practical success of many methods for nonparametric regression.
-One explanation that resolves this seeming paradox is the \emph{manifold hypothesis}, where the design points $X$ are assumed to lie on or near a manifold $\mc{X}$ of intrinsic dimension $m \ll d$.
+One explanation that resolves this seeming paradox is the \emph{manifold hypothesis}, where the design points $X$ are assumed to lie on or near a manifold $\mc{X}$ of intrinsic dimension $m < d$.
 Under the manifold hypothesis nonparametric regression is demonstrably easier, and in particular it is known \citep{bickel2007,ariascastro2018} that in this setting the optimal rates over $H^s(\Xset)$ scale like $n^{-2s/(2s + m)}$ (for estimation) and $n^{-4s/(4s + m)}$ (for testing).
@@
 We will also assume some regularity conditions on $\mc{X}$ and $P$.
-Recall that the injectivity radius of $\mc{X}$ is the maximum value of $\delta$ such that the exponential map $\exp_x: B_m(0,\delta) \subset T_x(\mc{X}) \to B_{\mc{X}}(x,\delta) \subset \mc{X}$ is a diffeomorphism for all $x \in \mc{X}$.
 \begin{enumerate}[label=(A\arabic*)]
@@
 \begin{itemize}
-	\item \textbf{Tuning parameters by sample-splitting.} 
-	\item \textbf{Other extensions.} We consider kernel smoothing because it is a simple and intuitive procedure that does not rely on knowledge of the domain $\mc{X}$ or distribution $P$.
+	\item \textbf{Tuning parameters by sample-splitting.} Extending the estimator $\wh{f}$ to be defined out-of-sample gives us the option of selecting the hyper-parameters $\varepsilon$ and $K$ in a data-dependent manner---e.g.
+by sample splitting, minimizing the error over a held-out set.
+This is by far the most common practical recommendation for selecting hyper-parameters \textcolor{red}{(references)}, and addresses a serious limitation of Laplacian eigenmaps.
+Moreover, it is known \textcolor{red}{(references)} that choosing hyper-parameters by minimizing error over a held-out set results in estimators that optimally adapt to the order of regularity $s$---that is, that are rate-optimal (up to $\log$ factors), even when $s$ is unknown.
+We believe that similar arguments will imply that $T_{h,n}\wh{f}$ is adaptive in this sense, though we do not pursue the details.
+	\item \textbf{Other extensions.} We consider extension by kernel smoothing because it is a simple and statistically optimal procedure that does not require any knowledge of the domain $\mc{X}$ or distribution $P$.
 There exist (many) other approaches to extending (or interpolating) a function $f$ based on the evaluations $\{f(X_1),\ldots,f(X_n)\}$, and we now comment on a few of these.
@@
 	\end{equation*}
-	A particularly popular version of this general approach is when $\mc{F}$ is an RKHS, which encompasses thin-plate spline interpolation $\mc{F} = H^s(\mc{X})$ for $s > d/2$ as a special case.
-This approach works well when $f$ is close to a function $u \in \mc{F}$ with reasonably small norm, which holds true by assumption  but as already discussed when $2s \leq d$ the Sobolev space $H^s(P)$ is not a Reproducing Kernel Hilbert Space.
-Another
-	\item \textbf{Comparison with kernel smoothing.} 
+	A particularly popular version of this general approach takes $\mc{F}$ to be an RKHS \textcolor{red}{(Rieger2008, Belkin2018)}, which encompasses thin-plate spline interpolation (where $\mc{F} = H^s(\mc{X})$ for $s > d/2$) as a special case.
+Naturally, this approach works well when $f$ is close to a function $u \in \mc{F}$ with reasonably small norm.
+This holds true when $f \in H^s(P;M)$ and $s > d/2$, but as already discussed when $s \leq d/2$ the Sobolev space $H^s(P)$ is not an RKHS, and in fact when $s \leq d/2$ thin-plate spline interpolation is ill-posed \citep{green93}.
+Another, arguably simpler approach is to extend $f$ to be piecewise constant on the Voronoi tessellation induced by $X_1,\ldots,X_n$, or equivalently to perform $1$-nearest neighbors regression on $f$.
+However, this approach is theoretically optimal only when $f \in C^1(P)$, in contrast to the kernel smoothing method we propose and study.
+	%Finally, the Nystr\"{o}m method extends  eigenvectors $v_k$ to functions $f_k: \mc{X} \to \Reals$ as follows,
+	%\begin{equation*}
+	%L_{n,\varepsilon}f_k = \lambda_k f_k~~\textrm{on $\mc{X}$, such that}~~f_k(X_i) = %v_k(X_i)~~\textrm{for $i = 1,\ldots,n$.}
+	%\end{equation*}
+	%Here $L_{n,\varepsilon}$ should be thought of as an operator acting on functions $f \in C(\mc{X})$ as follows,
+	%\begin{equation*}
+	%L_{n,\varepsilon}f(x) = \sum_{i = 1}^{n} (f(x) - f(X_i))\eta\Bigl(\frac{\|X_i - x\|}{\varepsilon}\Bigr)
+	%\end{equation*}
+	% The Nystr\"{o}m approach to extension makes intriguing (re-)use of the graph Laplacian $L_{n,\varepsilon}$ and its eigenvectors $v_k$.
+Similar methods have been analyzed in the context of kernel ridge regression (references), but little is known about its application to graph Laplacians.
 \end{itemize}
-
-\paragraph{Why not just kernel smoothing?}
-On the other hand, the allusion to kernel smoothing raises a very natural question: why should we prefer to first compute $\wh{f}$ and then smooth it out, as opposed to simply performing kernel smoothing over the original responses $Y$, that?
-
-We now answer this question, by \textcolor{red}{(...)}.
 
@@
 
-\textcolor{red}{(TODO)}: Come up with a set of experiments that might be interesting to run. Run them by Ryan and Siva, so that you can prioritize.
+\textcolor{red}{(TODO)}: Come up with a set of experiments that might be interesting to run.
+
+\begin{itemize}
+	\item \textbf{Laplacian eigenmaps approximates spectral series.} $1$d unit cube, uniform density, ``interesting'' regression function.
+Compare Laplacian eigenmaps with spectral series estimator, including their MSE curves, and their fits at several different values of $K$.
+	\item \textbf{Laplacian eigenmaps beats nonparametric least squares.} $1$d unit cube, very non-uniform density, same regression function.
+Make the same three comparisons, but in this case with respect to the least-squares estimator.
+	\item \textbf{Investigate our theoretical bounds on hyperparameters.} $1$d unit cube, uniform density, ``smooth'' regression function.
+Investigate the MSE heat map of Laplacian eigenmaps, as a function of $K$ and $\varepsilon$.
+	\item Swiss roll, density \textcolor{red}{(?)}, regression function \textcolor{red}{(?)}. Evaluate Laplacian eigenmaps in-sample error, and Laplacian-eigenmaps + kernel smoothing out-of-sample error.
+\end{itemize}
 
@@
 
-\textbf{(1) Future items:}
-\begin{itemize}
-	\item \textbf{Other properties of spectral series estimators.}
-\end{itemize}
+\subsection{Comparison with other estimators}
+In this paper, we have motivated Laplacian eigenmaps by viewing it as a noisy approximation of a classical spectral series method, which is its most obvious counterpart.
+We now discuss the relationship between Laplacian eigenmaps and three other approaches to nonparametric regression.
+The first two---nonparametric least squares and kernel smoothing---are classical, whereas the third---Laplacian smoothing---makes use of the graph Laplacian in a different way than Laplacian eigenmaps.
+
+\paragraph{Least squares.}
+The standard recommended alternative to spectral series methods, when the distribution $P$ is considered non-uniform or unknown, is to do least-squares.
+Concretely, suppose instead we had access only to eigenfunctions $f_1',f_2',\ldots$ of the unweighted Laplace-Beltrami operator $\Delta = \Delta_{\nu}$.
+Then letting $F_K = \mathrm{span}\{f_1,\ldots,f_K\}$, the least-squares estimator and test statistic
+\begin{equation*}
+\wt{f}_{\mathrm{LS}} := \argmin_{f \in F_K} \|Y - f\|_n^2,\textrm{and}~~\wt{T}_{\mathrm{LS}} = \|\wt{f}\|_{\nu}^2
+\end{equation*}
+are still rate-optimal over $H_0^s(P)$.
+However, this is not a totally satisfactory fix.
+For one thing, the least squares approach still requires that we know $\Delta$, and is not well suited for the case where $\mc{X}$ is a manifold, and $\Delta$ is the manifold Laplace-Beltrami operator.
+Additionally, diagonalizing even the unweighted Laplace-Beltrami operator $\Delta$ is quite difficult for all but a few special domains, such as $\mc{X} = [0,1]^d$.
+
+\paragraph{Kernel smoothing.}
+It is also natural to ask whether the two stage estimator $T_{h,n}\wh{f}$ defined in Section~\ref{sec:out_of_sample} has any advantage over the simpler approach of directly kernel smoothing the responses, i.e.
+using the estimator $T_{h,n}Y$ (possibly for a different choice of $h$).
+In Appendix~\textcolor{red}{(?)}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\|f_0 - T_{h,n}\wh{f}\|_P^2 \ll \inf_{h'} \|f_0 - T_{h',n}Y\|_P^2$.
+This is possible because Laplacian eigenmaps induces a completely different bias than kernel smoothing.
+In particular, when $f_0$ and $p$ satisfy the so-called \emph{cluster} assumption--- i.e.
+$f_0$ is piecewise constant in high-density regions of $p$--- then the bias of Laplacian eigenmaps can much smaller than that of kernel smoothing (for equivalent levels of variance).
+
+We emphasize that this does not contradict the well-known fact that kernel smoothing is an optimal method for nonparametric regression over e.g.
+H\"{o}lder balls.
+It simply reflects that in the standard nonparametric regression setup---which we adopt in the main part of this paper, and in which $P$ is assumed to be equivalent to Lebesgue measure---the biases of Laplacian eigenmaps and kernel smoothing are equivalent.
+There has been some work \textcolor{red}{(Rigollet, Wasserman, Niyogi, El Alaoui)} analyzing semi-supervised learning in the case where $f_0$ and $p$ satisfy some special relationship, such as the cluster assumption.
+However, a comprehensive analysis comparing the error of Laplacian based methods (e.g.
+eigenmaps) to that of other methods remains outstanding.
+
+\subsection{Future Work}
+As mentioned at the outset, our work can be viewed as a contribution both to the fields of nonparametric regression with series estimators, and to graph-based learning.
+Fitting, we end our discussion by mentioning some open work in each of these directions.
+
+Much is known about classical spectral series methods beyond their rate optimality.
+For instance, such estimators and tests exhibit \emph{sharp optimality}, meaning their risk can be bounded to within a $(1 + o(1))$ factor of the optimal risk.
+They can adapt to unknown smoothness of the regression function.
+They can be used to estimate smooth functionals of the regression function, or alternatively to form confidence sets in $L^2(P)$.
+It would be interesting to see if Laplacian eigenmaps could replicate the performance of classical methods in any, or all, of these problems.
+
+On the other hand, there are many variants of Laplacian eigenmaps worth considering.
+For instance, one can change the graph under consideration (e.g.
+by using the k-nearest neighbors), or the normalization of the graph Laplacian $L_{n,\varepsilon}$ (e.g.
+by using the symmetric normalized Laplacian).
+The former is practically useful, because it typically leads to connected graphs while always ensuring a given level of edge sparsity.
+In the latter, the graph Laplacian converges to a different limiting operator, which possesses different eigenvectors than $\Delta_P$ and thereby induces a different bias.
+We believe that under the setup we consider here, both methods will continue to be optimal.
 

### 2021-03-17 notes/misc/insights_and_connections.tex

@@
 	\end{enumerate}
+	In a negative direction,~\eqref{eqn:bandlimited_laplacian_smoothing} reveals that the Laplacian smoothing is automatically zeroing out certain high-frequency signals.
+Once we acknowledge this fact, we should probably also acknowledge that the particular choice of frequency at which to cut off was made somewhat arbitrarily.
+Indeed, we get much stronger theoretical guarantees if we cut off at a much lower frequency.
+This motivates Laplacian eigenmaps.
 \end{itemize}

### 2021-03-17 talks/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
7/11 slides of first draft of Laplacian smoothing paper done.

\documentclass{beamer}

% Packages
\usepackage{bm}

%% Theme related.
\usetheme{default}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{frametitle}[default][center]
\setbeamertemplate{itemize items}[circle]
\usefonttheme[onlymath]{serif}

%% Math
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\wt}[1]{\widetilde{#1}}
\DeclareMathOperator*{\argmin}{argmin}

%% Font
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}

\title{Laplacian Smoothing on Neighborhood Graphs}

\author{Alden Green}
\institute{Department of Statistics and Data Science \\
		   Carnegie Mellon Unviersity}
\date{}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}{Basic problem setup}
\alert{Nonparametric regression, random design}: We observe $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are sampled independently from $P$, and
\begin{equation*}
Y_i = f_0(X_i) + \varepsilon_i, \quad \varepsilon_i \sim N(0,1).
\end{equation*}
Goal is to learn the unknown regression function $f_0$, which is assumed to be \alert{Sobolev} smooth, i.e.
\begin{equation*}
\|f_0\|_{H^1}^2 := \int \|\nabla f_0(x)\|^2 \,dx \quad \text{is small.}
\end{equation*}
Either \alert{estimate} $f_0$ with some $\hat{f}$, or \alert{test} whether $f_0 = 0$.
\end{frame}

\begin{frame}{Laplacian smoothing}
Laplacian smoothing solves a discrete, penalized least squares problem. 
\begin{enumerate}
	\item  Form a \alert{neighborhood graph} $G$ over $X_1,\ldots,X_n$, with weighted edges $W_{i,j} = K(\|X_i - X_j\|/r)$. The \alert{graph Laplacian} $L$ is defined by
	\begin{equation*}
	(Lf)_i = \sum_{i = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr) W_{i,j}.
	\end{equation*}
	\item The \alert{Laplacian smoothing} estimator is
	\begin{equation*}
	\wh{f} = \argmin_{f \in \Reals^n} \|Y - f\|_n^2 + \frac{\lambda}{2} f^{\top} L f, \end{equation*}
	with corresponding test statistic $\wh{T} = \|\wh{f}\|_n^2$.
\end{enumerate}
\textcolor{red}{(TODO): Simple picture here.}
\end{frame}

\begin{frame}{Motivation}
Why Laplacian smoothing?
\begin{itemize}
	\item \emph{Computational ease.} Sparse graph, fast solvers.
	\item \emph{Generality.} Makes sense whenever one can form a graph $G$. 
	\item \emph{Weak supervision.} Naturally extends to semi- and unsupervised problems.
\end{itemize}
Intuition: when $n$ large, $r$ small and $f$ smooth, \textcolor{blue}{(Bousquet et al.)} show
\begin{equation*}
f^T L f = \frac{1}{2}\sum_{i,j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 W_{i,j} \approx n^2 r^{d + 2} \int \|\nabla f\|_2^2 \,dx.
\end{equation*}
Seemingly $\wh{f}$ is a noisy approximation of a very classical estimator $\wt{f}$: aka \alert{smoothing spline} (1d), or \alert{thin-plate spline} ($d > 1$), defined as
\begin{equation*}
\widetilde{f} = \argmin_{f \in H^1} \|Y - f\|_n^2 + \lambda \int \|\nabla f(x)\|_2^2 \,dx.
\end{equation*}
Similarly $\wh{T}$ approximates $\wt{T} = \|\wt{f}\|_n^2$. 
\end{frame}

\begin{frame}{Main Question}
In $1d$, smoothing splines achieve the minimax optimal rate \blue{$n^{-2/3}$} (estimation) or \blue{$n^{-4/5}$} (testing).
\begin{block}{{\bf Question}}
	Does Laplacian smoothing enjoy the same optimality properties as smoothing/thin plate splines?
\end{block}

\begin{block}{{\bf Answer}}
	They actually do better! When properly tuned:
	\begin{itemize}
		\item Laplacian smoothing has error of \blue{$n^{-2/3}$} (estimation) and \blue{$n^{-4/5}$} (testing) when $d = 1$, matching smoothing splines.
		\item Laplacian smoothing also achieves optimal error \blue{$n^{-2/(2 + d)}$} (estimation) and \blue{$n^{-4/(4 + d)}$} (testing) when $d = 2,3,4$ (up to $\log(n)$ factors when $d = 4$).
		\item On the other hand $\wt{f}$ is \alert{not even well-defined} when $d > 1$. Well known in spline theory community.
	\end{itemize}
\end{block}
\end{frame}

\begin{frame}{Summary of Main Results}

{\large Estimation rates, assuming $f_0 \in H^1(M)$}
\begin{table}
	\begin{center}
		\begin{tabular}{p{.25\textwidth} | p{.25\textwidth} p{.25\textwidth} }
			Dimension & Laplacian smoothing & Thin-plate splines \\
			\hline
			$d = 1$ & \blue{${n^{-2/3}}$} & \blue{${n^{-2/3}}$} \\
			$d = 2,3$ & \blue{${n^{-2/(2 + d)}}$} & \red{NA} \\
			$d = 4$ & $\blue{{n^{-1/3}}} (\log n)^{1/3}$ & \red{NA} \\
			$d \geq 5$  & $(\log n/n)^{4/(3d)}$ & \red{NA} \\
		\end{tabular}
	\end{center}
\end{table}

{\large Testing rates, assuming $f_0 \in H^1(M)$ ($d < 4$) or $f_0 \in L^4(M)$ ($d \geq 4$)}
\begin{table}
	\begin{center}
		\begin{tabular}{p{.25\textwidth} | p{.25\textwidth} p{.25\textwidth} }
			Dimension & Laplacian smoothing & Thin-plate splines \\
			\hline
			$d = 1$ & \blue{${n^{-4/5}}$} & \blue{${n^{-4/5}}$} \\
			$d = 2,3$ & \blue{${n^{-4/(4 + d)}}$} & \red{NA} \\
			$d \geq 4$ & $\blue{{n^{-1/2}}}$ & \red{NA} \\
		\end{tabular}
	\end{center}
\end{table}
\end{frame}

\begin{frame}{Explanation: Failure of Thin-plate Splines}
Well-posedness of smoothing splines (when $d = 1$) vs. ill-posedness of thin-plate splines ($d \geq 2$) explained by \alert{Sobolev Embedding Theorem}.
\begin{itemize}
	\item In $1d$, every function in $H^1$ is continuous. 
	\item When $d \geq 2$, this is no longer the case.
\end{itemize}
As a result the criterion underlying $\wt{f}$ is \red{not continuous} with respect to $H^1$ norm. I can construct sequences of ``bump functions'' $f_1,f_2,\ldots$ such that
\begin{equation*}
f_k(X_i) = Y_i \quad \text{for all $k$, but} \quad \|f_k\|_{H^1} \to 0.
\end{equation*}
Implication: For any $\delta > 0$, there exists a function $f$ which satisfies
\begin{equation*}
f(X_i) = Y_i \quad \text{for all $i = 1,\ldots,n$, yet} \int\|\nabla f(x)\|^2 \,dx < \delta. 
\end{equation*} 
\end{frame}

\begin{frame}{Explanation: Success of Laplacian Smoothing}
Roughly, discretization is a blessing.

\end{frame}

\begin{frame}{What about $d > 4$?}
content...
\end{frame}

\begin{frame}{Extensions}

\end{frame}

\begin{frame}{Analysis}
\end{frame}

\end{document}

### 2021-03-18 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Finished editting higher-order section of Laplacian eigenmaps paper.

@@
 \end{equation}
-The Laplacian $L$ can be seen as a discretization of the Laplace-Beltrami operator $\Delta_P$ \textcolor{red}{(...)}.
+The Laplacian $L$ can be seen as a discretization of the Laplace-Beltrami operator $\Delta_P$, defined as
+\begin{equation}
+\label{eqn:fokker_planck_1}
+\Delta_Pf= \frac{-1}{p} \mathrm{div}(p^2 \nabla f) = \Delta f - 2\frac{\nabla p^{\top} \nabla f}{p}.
+\end{equation}
 The eigenvectors of $L$ form an orthonormal basis of $L^2(X)$, and are approximations to the eigenvectors of $\Delta_P$.
@@
 			\multirow{3}{*}{$s = 2$ or $3$} & $\dim(\mc{X}) \leq 4$  & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
-			& $4 <\dim(\mc{X}) < 4s$  & \textcolor{red}{(?)}& \textcolor{red}{(?)} \\
+			& $4 <\dim(\mc{X}) < 4s$  & $n^{-2s/(2(s - 1) + d)}$ & $n^{-2s/(2(s - 1) + m)}$\\
 			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
@@
 	\item \textcolor{red}{(TODO)} Summarize what is known on fixed graphs.
-	\item \textcolor{red}{von Luxburg, Belkin, Slepcev} show that for a fixed index $k$, the eigenvector $v_k$ converges to the eigenfunction $u_k$, in the sense that \textcolor{red}{(...)}.
+	\item \textcolor{red}{von Luxburg, Belkin, Slepcev} show that for a fixed index $k$, the eigenvector $v_k$ converges to the eigenfunction $\psi_k$, in the sense that \textcolor{red}{(...)}.
 \textcolor{red}{Shi, Burago, Garcia Trillos, Calder, Cheng} build on this, giving finite sample bounds, rates of convergence, and making statements uniform over $1 \leq k \leq k_{\max}$.
@@
 	\item Define asymptotic notation: $\ll$, $\lesssim$ and $\asymp$.
+	\item Define min and max notation: $\vee$ and $\wedge$.
 \end{itemize}
@@
 As a result, is natural that they accurately approximate only those functions $f_0 \in \mc{H}^s(P) \subset H^s(P)$.
-Although the zero-trace boundary condition is more restrictive than~\eqref{eqn:condition}---since it also requires that the even derivatives be equal to $0$---the point is that some kind of boundary condition will be necessary in order to obtain optimal rates for spectral series methods, including Laplacian eigenmaps.
-We will stick to the zero-trace condition, which greatly eases some steps in our proofs.
+Although the zero-trace boundary condition is more restrictive than~\eqref{eqn:condition}---since it also requires that derivatives of even-order be equal to $0$---the point is that some kind of boundary condition will be necessary in order to obtain optimal rates for spectral series methods, including Laplacian eigenmaps.
+We will stick to the zero-trace condition, since it greatly eases some of the steps in our proofs.
 
@@
 Tsybakov,Johnstone,Gine)}, it can be shown the estimator
-\begin{equation*}
+\begin{equation}
+\label{eqn:spectral_series_estimator}
 \wt{f} := \sum_{k = 1}^{K} \dotp{Y}{\psi_k}_n \psi_k,
-\end{equation*}
+\end{equation}
 achieves the minimax rate over $\mc{H}^s(P;M)$, and the same upper bound on the minimax rate therefore applies to $H_0^s(P;M)$.
@@
 Thus the minimax rates are unchanged---we call estimators which achieve these optimal rates \emph{in-sample minimax rate-optimal estimators}.
+
+There is one other subtlety introduced by the use of in-sample mean squared error.
+Technically speaking, elements $f \in H^s(P)$ are equivalence classes, defined only up to a set of measure zero.
+Thus one cannot speak of the pointwise evaluation $f_0(X_i)$, as we do by defining our target of estimation to be $f_0(X_i)$, $i=1,\ldots,n$, until one selects \emph{representatives}.
+When $s > d/2$, every element $f$ of $H^s(P)$ admits a continuous version $f^{\ast}$, and it is standard to pick this as the favored representative.
+When $s \leq d/2$, some elements in $H^s(P)$ do not have any continuous version; however they admit a \emph{quasi-continuous} version \citep{evans15} known as the \emph{precise representative}, and we use this representative.
+To be clear, however, it does not really matter which representative we choose.
+Since all versions agree except on a set of measure zero, and since $P$ is absolutely continuous with respect to Lebesgue measure (in Model~\ref{def:model_flat_euclidean}) or the volume form $d\mu$ (in Model~\ref{def:model_manifold}), with probability $1$ any two versions $g_0, h_0 \in f_0$ will satisfy $g_0(X_i) = h_0(X_i)$ for all $i = 1,\ldots,n$.
+The bottom line is that we can use the notation $f_0(X_i)$ without fear of ambiguity or confusion.
 
@@
 
-\paragraph{Estimation.} When the graph radius $\varepsilon$ and number of eigenvectors $K$ are chosen appropriately, we show in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that the estimator $\wh{f}$ achieves the minimax rate, when error is measured in $\|\cdot\|_{n}$.
+\paragraph{Estimation.} When the graph radius $\varepsilon$ and number of eigenvectors $K$ are chosen appropriately, we show in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that the estimator $\wh{f}$ achieves the minimax rate over $H^1(P;M)$, when error is measured in $\|\cdot\|_{n}$.
 \begin{enumerate}[label=(A\arabic*)]
-	\setcounter{enumi}{3}
+	\setcounter{enumi}{0}
 	\item 
@@
 On the other hand, when $M \gg n^{1/d}$, then computing Laplacian eigenmaps with $K = n$ achieves the rate $\|\wh{f} - f_0\|_n^2 \asymp 1$, which is better than the rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}.
-However, in truth these are edge cases, \textcolor{red}{(...)}.
+However, in truth these are edge cases, which do not fall neatly into the framework of nonparametric regression.
 	\item The lower bound on $\varepsilon$ imposed by~\eqref{eqn:radius_fo} is on the order of the connectivity threshold, the smallest length scale at which the resulting graph will still, with high probability, be connected.
@@
 \begin{equation*}
-t_{a} := \frac{K}{n} + \frac{1}{n}\sqrt{\frac{2K}{a}}
+t_{a} := \frac{K}{n} + \frac{1}{n}\sqrt{\frac{2K}{a}}.
 \end{equation*}
@@
 \begin{enumerate}[label=(A\arabic*)]
-	\setcounter{enumi}{4}
+	\setcounter{enumi}{1}
 	\item 
@@
 	\label{eqn:eigenvector_testing_fo}
-	K \asymp \min\Bigl\{(M^2n)^{2d/(4 + d)} \vee 1, n\Bigr\}
+	K \asymp \min\Bigl\{(M^2n)^{2d/(4 + d)} \vee 1, n\Bigr\}.
 	\end{equation}
@@
 Suppose additionally $f_0 \in H^1(P,M)$, and that $d < 4$.
-Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if $f_0$ satisfies
+Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_testing_fo}, and if $f_0$ satisfies
 	\begin{equation}
 	\label{eqn:laplacian_eigenmaps_testing_criticalradius_fo}
-	\|f_0\|_n^2 \geq \frac{C}{\delta}\biggl(\Bigl(M^2(M^2n)^{-4/(4 + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{1}{\delta n^{p(d)}} \biggr) \vee \frac{1}{n}
+	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-4/(4 + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{p(d)}} \biggr) \vee \frac{1}{n}
 	\end{equation}
@@
 \end{theorem}
-Although~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} involves taking the maximum of several different terms, the most important takeaway is that if $n^{-1/2} \lesssim M^2 \ll n^{1/4}$ ($d = 1$) or $n^{-1/2} \lesssim M^2 \ll n^{2/d - 1/2}$ ($d = 2$ or $d = 3$), then $\varphi$ is a minimax rate-optimal test over $H^1(\mc{X},M)$.
+Although~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} involves taking the maximum of several different terms, the important takeaway of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} is that if $n^{-1/2} \lesssim M^2 \ll n^{1/4}$ ($d = 1$) or $n^{-1/2} \lesssim M^2 \ll n^{2/d - 1/2}$ ($d = 2$ or $d = 3$), then $\varphi$ is a minimax rate-optimal test over $H^1(P,M)$.
 
@@
 	\item When $d \geq 4$, as we have already mentioned $H^1(\mc{X})$ does not continuously embed into $\Leb^4(\mc{X})$.
-This has severe consequences for test statistics which involve taking an $L^2(P_n)$-norm, such as $\wh{T}$, or $L^2(P)$ norm, such as $\wt{T}$, since they may no longer have a finite variance.
+This has severe consequences for test statistics based on the $L^2(P_n)$-norm---such as $\wh{T}$---or $L^2(P)$ norm---such as $\wt{T}$---since they may no longer have a finite variance.
 Indeed, as far we know when $d \geq 4$ the optimal testing rates over $H^1(\mc{X})$ are not known.
 On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then we show in Appendix~\textcolor{red}{(?)} that the Laplacian eigenmaps test, with $K = n$, is minimax optimal whenever $\|f_0\|_n^2 \gtrsim n^{-1/2}$.
-	\item \textcolor{red}{(TODO)}: Explain the presence of the $n^{p(d)}$ term. Explain that if you assume $f_0 \in C^1(\mc{X};M)$, you can avoid this term.
-	\item \textcolor{red}{(TODO)}: Explain what else you can get when you assume $C^1(\mc{X};M)$.
+	\item The requirement that $\|f_0\|_P^2 \gtrsim n^{-p(d)}$ is necessary in order for us to establish that with probability at least $1 - 5\delta$,
+	\begin{equation}
+	\label{eqn:small_ball_estimate}
+	\|f_0\|_n^2 \geq \frac{1}{\delta}\|f_0\|_P^2.
+	\end{equation}
+	This requirement prevents us from showing that $\varphi$ is a minimax optimal test when $M \gtrsim n^{1/4}$ ($d = 1$) or $M \gtrsim n^{2/d - 1/2}$ ($d = 2$ or $d = 3$).
+Of course, note that we do show $\varphi$ is an optimal test when $M \asymp 1$, which is by far the most commonly considered case.
+On the other hand, if we assume $f_0 \in C^1(\mc{X};M)$, then we can establish~\eqref{eqn:small_ball_estimate}, and indeed stronger estimates, without any restrictions on $\|f_0\|_P^2$.
+As a result, under this stronger assumption we can show that $\varphi$ is an optimal test for all $n^{-1/2} \lesssim M^2 \lesssim n^{1/d}$, matching our results in the estimation case.
 \end{itemize}
@@
 \label{sec:higher_order_sobolev_classes}
-In this section, we now assume that the regression function displays some higher-order regularity and is zero-trace, i.e.
+We now assume that the regression function displays some higher-order regularity and is zero-trace, i.e.
 $f_0 \in H_0^s(P)$.
-We point out that on a graph there is no natural analogue to higher-order derivatives.
-For this reason it is not at all obvious that graph-based methods \emph{can} adapt to higher-order smoothness.
-Remarkably, we show that this is in fact the case.
-Under appropriate choices of $\varepsilon$ and $K$, the estimator $\wh{f}$ has (in-sample) optimal error (up to constant factors) $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2s/(2s + d)}$.
-
+At first glance, it is not at all clear that Laplacian eigenmaps \emph{can} take advantage of this additional regularity, and it is especially unclear whether it can do so in an optimal manner.
+As $s \to \infty$ and $d$ remains fixed, the optimal estimation and testing rates approach the parametric rate $n^{-1}$.
+On the other hand, the sharpest known results show that Laplacian eigenvectors $v_k$ converge to continuum eigenfunctions $\psi_k$ only at a rate of \textcolor{red}{(?)}, which is obviously much slower than $n^{-1}$.
+
+Nevertheless, as we will establish momentarily, remarkably Laplacian eigenmaps methods continue to be optimal for all orders of $s$, as long as the design density is itself also sufficiently regular, $p \in C^{s - 1}(\mc{X})$.
+In estimation, this is the case for any dimension $d$, whereas in testing it is the case only when $d \leq 4$.
+
+\paragraph{Estimation.}
+In order to show that $\wh{f}$ is an optimal estimator over $H_0^s(P;M)$, we will require that $\varepsilon$ be meaningfully larger than the lower bound in~\ref{asmp:parameters_estimation_fo}.
 \begin{enumerate}[label=(A\arabic*)]
-	\setcounter{enumi}{5}
+	\setcounter{enumi}{2}
 	\item 
@@
 	The graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy
-	\begin{equation*}
-	C_0 \cdot \max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/d}, n^{-1/(2r + d)}\biggr\} \leq \varepsilon \leq \min\{i_0, K^{-1/d}\}
-	\end{equation*}
+	\begin{equation}
+	\label{eqn:radius_ho}
+	\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/d}, (M^2n)^{-1/(2(s - 1) + d)}\biggr\} \ll \varepsilon \ll \min\{1, K^{-1/d}\}
+	\end{equation}
 	and
 	\begin{equation*}
-	K \asymp (M^2n)^{2d/(2s + d)}
+	K \asymp (M^2n)^{2d/(2s + d)}.
 	\end{equation*}
 \end{enumerate}
-
+It is important to note that the two restrictions of~\ref{asmp:parameters_estimation_ho} are not mutually exclusive.
+So long as $M^2 \gg n^{-1}$, then $(M^2n)^{-2/(2(s - 1) + d)} \ll (M^2n)^{-2/(2s + d)}$ for all values of $s$ and $d$.
 \begin{theorem}
 	\label{thm:laplacian_eigenmaps_estimation_ho}
-	Under appropriate conditions, with probability at least $1 - \delta$,
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H_0^s(P,M)$ and $p \in C^{s - 1}(\mc{X})$.
+There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_ho}, then
 	\begin{equation}
 	\label{eqn:laplacian_eigenmaps_estimation_ho}
-	\|\wh{f} - f_0\|_n^2 \lesssim n^{-2s/(2s + d)}
-	\end{equation}
+	\|\wh{f} - f_0\|_n^2 \leq C\Bigl(\frac{1}{\delta}M^2(M^2n)^{-2s/(2s + d)} \wedge 1\Bigr) \vee \frac{1}{n},
+	\end{equation}
+	with probability at least $1 - \delta - Cn\exp(-cn\varepsilon^d)$.
 \end{theorem}
-Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that in the flat Euclidean setting Laplacian eigenmaps is a rate-optimal estimator over Sobolev classes, for all scalings of $s$ and $d$.
+Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that in the flat Euclidean setting Laplacian eigenmaps is a rate-optimal estimator over Sobolev classes, for all values of $s$ and $d$.
 Some other remarks:
 \begin{itemize}
-	\item \textbf{(1.1) Optimal rates, no RKHS.} It is worth pointing out that we do not require that the regularity of the Sobolev space satisfy $2s > d$, a condition often seen in the literature.
-In the ``sub-critical'' regime $2s < d$, the Sobolev space $H^s(\mc{X})$ is quite irregular.
-For instance, it is not a Reproducing Kernel Hilbert Space (RKHS), nor does it continuously embed into $C^0(\mc{X})$, much less into any H\"{o}lder space.
-As a result, certain versions of the nonparametric regression problem are ill-posed---for instance when loss is measured in $\Leb^{\infty}$ norm, or when the design points $\{X_1,\ldots,X_n\}$ are assumed to be fixed.
-Likewise, certain estimators are ``off the table'', most notably RKHS-based methods such as smoothing splines.
-Nevertheless, for random design regression with error measured in $\Leb^2(P)$-norm, spectral series estimators obtain the minimax rates for all values of $s$ and $d$.
-Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} shows that the same is true with respect to Laplacian eigenmaps, with error measured in $\Leb^2(P_n)$ norm.
-	\item \textbf{(1.2) The zero-trace condition.} The zero-trace condition can likely be weakened.
-However, requiring some boundary condition on $f_0$ and its first through $(k -1 )$th-order derivatives is unavoidable.
-\textcolor{red}{Explain why: graphs are implicitly boundaryless objects, and the eigenvectors of $L$ converge to those eigenvectors of $\Delta_P$ which satisfy boundary conditions.
-Repeat that this is a standard assumption made when analyzing spectral series estimators}.
-	\item \textbf{(1.3) The bounded norm and smooth density conditions.} We require that $f_0$ be bounded in the Sobolev norm $\|f_0\|_{H^s(\mc{X})}$---as opposed to the semi-norm $|f_0|_{H^s(\mc{X})}$--- and that $p \in C^{s - 1}(\mc{X})$.
-Both requirements are related and fundamental, as we shall now explain.
-As previously mentioned, $L$ converges to a weighted-version of the Laplace operator $\Delta_P$ (also known as the Fokker-Planck) operator.
-The operator $\Delta_P$ includes both first and second-order derivatives,
-	\begin{equation}
-	\label{eqn:fokker_planck_1}
-	\Delta_Pf= \frac{-1}{p} \mathrm{div}(p^2 \nabla f) = \Delta f - 2\frac{\nabla p^{\top} \nabla f}{p},
-	\end{equation}
-	and therefore a function $f \in H_0^s(\mc{X})$ has finite energy $\|\Delta_P f\|_{P}^2$ only if the first-order partial derivatives of $f$ and $p$ are bounded.
-\textcolor{red}{This discussion---and in particular the last sentence---are obviously not quite correct.
-This is in part because you realize now that the requirements are not in fact fundamental.
-Instead, the requirement should merely be that $\|\Delta_P^kf\|_{\Leb^2(P)}^2$ is bounded---here $k = s/2$ and we assume $s$ is even---and the upper bounds should be in terms of this semi-norm.}.
-	\item \textbf{(1.4) Both of the remarks after Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} continue to apply.}
+	\item We do not require that the regularity of the Sobolev space satisfy $s > d/2$, a condition often seen in the literature.
+In the sub-critical regime $s \leq d/2$, the Sobolev space $H^s(\mc{X})$ is quite irregular.
+It is not a Reproducing Kernel Hilbert Space (RKHS), nor does it continuously embed into $C^0(\mc{X})$, much less into any H\"{o}lder space.
+As a result, for certain versions of the nonparametric regression problem---e.g.
+when loss is measured in $\Leb^{\infty}$ norm, or when the design points $\{X_1,\ldots,X_n\}$ are assumed to be fixed---in a minimax sense even consistent estimation is not possible.
+Likewise, certain estimators are ``off the table'', most notably RKHS-based methods such as thin-plate splines of degree $k \leq d/2$.
+Nevertheless, for random design regression with error measured in $\Leb^2(P)$-norm, the spectral series estimator $\wt{f}$ defined in~\eqref{eqn:spectral_series_estimator} obtains the ``usual'' minimax rates $n^{-2s/(2s + d)}$ for all values of $s$ and $d$.
+Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and \ref{thm:laplacian_eigenmaps_estimation_ho} show that the same is true with respect to Laplacian eigenmaps, with error measured in $\Leb^2(P_n)$-norm.
+	\item \textcolor{red}{(TODO 1): This entire comment is too vague right now.
+I'll leave it in, in case Siva + Ryan think it is sufficiently valuable to edit and make more precise.} The lower bound on $\varepsilon$ means that the graph $G$ will be quite dense, with average degree growing polynomially in $n$ as $n \to \infty$.
+Consequently, exactly computing the eigenvectors $v_k$ is more computationally intensive than if the graph radius were on the order of the connectivity threshold $\varepsilon \asymp (\log n/n)^{1/d}$.
+Recall that in the first-order case, we established in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that this smaller choice of $\varepsilon$ still resulted in an optimal estimator.
+Thus Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and~\ref{thm:laplacian_eigenmaps_estimation_ho} can be viewed as revealing a tradeoff between statistical and computational efficiency; although to be clear, we have no theoretical evidence that Laplacian eigenmaps \emph{fails} to adapt to higher-order smoothness when $\varepsilon \asymp (\log n/n)^{1/d}$---we simply cannot prove that it succeeds.
+	\textcolor{red}{(TODO 2): If we decide it is worth our time to investigate the optimal choice of graph radius empirically, we can add a line here mentioning this.}
+
+	Suppose one does choose $\varepsilon$ meaningfully larger than the connectivity threshold, as our theory requires when $s > 1$.
+We now discuss a procedure to efficiently compute an approximation to the Laplacian eigenmaps estimate, without meaningfully increasing the statistical error incurred: \emph{edge sparsification}.
+For a given $\delta \in (0,1)$, various algorithms have been designed \textcolor{red}{(references)} to efficiently remove many edges from the graph $G$, resulting in a sparser graph $\wc{G}$ with Laplacian $\wc{L}_{n,\varepsilon}$, while only perturbing the spectrum of the Laplacian by a factor of at most $\sigma \geq 1$, meaning
+	\begin{equation*}
+	\frac{1}{\sigma} \cdot u^{\top} \wc{L}_{n,\varepsilon} u \leq u^{\top} L_{n,\varepsilon} u \leq \sigma \cdot u^{\top} \wc{L}_{n,\varepsilon}u \quad \textrm{for all $u \in \Reals^n$.}
+	\end{equation*}
+	The graph $\wc{G}$ is said to be $\sigma$-spectrally similar to $G$.
+Let $\wc{f}$ be the Laplacian eigenmaps estimator, computed with respect to the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ .
+Because $\wc{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wc{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$ \textcolor{red}{(reference needed)}.
+Statistically speaking, as we show in Appendix~\textcolor{red}{(?)} $\wc{f}$ has $L^2(P_n)$-error of at most $\sigma^{2s}$ times the error of $\wh{f}$.
+Therefore, as long as $\sigma \asymp 1$ the estimator $\wc{f}$ will also be rate-optimal.
+Indeed, as pointed out by~\citet{sadhanala16b}, geometric graphs such as $G$ exhibit a special structure.
+In this case there exist far simpler and faster methods for sparsification than those of~\textcolor{red}{(references)}, which, at least empirically, appear to suffice.
+	\item The requirement $p \in C^{s - 1}(\mc{X})$ is a strong condition, but is essential to showing that $\wh{f}$ enjoys faster rates of convergence.
+We give more detail about this when we discuss our analysis in Section~\ref{subsec:analysis}.
 \end{itemize}
 
-\textbf{(1.4) Rates for testing with higher-order smoothness assumptions.} For testing, we can get the optimal rate $n^{-4s/(4s + d)}$ when $d \leq 4$.
-When $4 < d \leq 4s$, we get the rate~\textcolor{red}{(...)} Recall that the rate $n^{-4s/(4s + d)}$ is optimal whenever $d \leq 4s$.
-Thus when $4 < d < 4s$, our upper bounds do not imply that $\varphi$ is an optimal test.
-We do not know whether this is due to looseness in our proof techniques, or a problem with the test itself, and leave this for future work.
+\paragraph{Testing.} The test $\varphi$ can adapt to the higher-order smoothness of $f_0$, when $\varepsilon$ and $K$ are chosen correctly.
+\begin{enumerate}[label=(A\arabic*)]
+	\setcounter{enumi}{3}
+	\item 
+	\label{asmp:parameters_testing_ho}
+	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy~\eqref{eqn:radius_ho}. Additionally,
+	\begin{equation}
+	\label{eqn:eigenvector_testing_ho}
+	K \asymp \min\Bigl\{(M^2n)^{2d/(4s + d)} \vee 1, n\Bigr\}.
+	\end{equation}
+\end{enumerate}
+When $d \leq 4$, for any value of $s$ it is possible to choose $\varepsilon$ and $K$ such that both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho} are satisfied, and our next theorem establishes that $\varphi$ is an optimal test.
+In what follows, let $p(d) = \min\{1,2s/d\}$.
+\begin{theorem}
+	\label{thm:laplacian_eigenmaps_testing_ho}
+	Fix $a,b \in (0,1)$.
+Suppose Model~\ref{def:model_flat_euclidean}.
+Then $\mathbb{E}_0[\varphi] \leq a$, i.e $\varphi$ is a level-$a$ test.
+Suppose additionally $f_0 \in H_0^s(P,M)$,that $p \in C^{s-1}(\mc{X})$, and that $d \leq 4$.
+Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_testing_ho}, and if $f_0$ satisfies
+	\begin{equation}
+	\label{eqn:laplacian_eigenmaps_testing_criticalradius_ho}
+	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-4s/(4s + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{p(d)}} \biggr) \vee \frac{1}{n},
+	\end{equation}
+	then $\Ebb_{f_0}[1 - \phi] \leq b$.
+\end{theorem}
+Just as in the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that $\varphi$ is a minimax optimal test over $H_0^s(P)$ when $n^{-1/2} \ll M^2 \ll n^{1/4}$ (if $2s < d$), or if $n^{-1/2} \ll M^2 \ll (2/d - 1/2)$ (if $2s \geq d$.)
+
+However, unlike the first-order case, when $4 < d < 4s$ the optimal testing rate over $H_0^s(P)$ continues to be $n^{-4s/(4s + d)}$.
+Unfortunately, we cannot claim that $\varphi$ is an optimal test in this regime.
+The culprit is that there no longer exist any choices of $\varepsilon,K$ which satisfy both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho}, and we can therefore no longer balance testing bias and variance (defined momentarily in~\textcolor{red}{(?)}).
+From Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, we can gather that $\varphi$ must have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2s + d)}$.
+In fact, we can derive slightly better (though still not optimal) rates.
+\begin{theorem}
+	\label{thm:laplacian_eigenmaps_testing_ho_suboptimal}
+	Under the same setup as Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, but with $d > 4$.
+If the Laplacian eigenmaps test $\varphi$ is computed with $K$ satisfying~\eqref{eqn:eigenvector_testing_ho}, and $\varepsilon \asymp K^{-1/d}$, and if
+	\begin{equation}
+	\label{eqn:laplacian_eigenmaps_testing_criticalradius_ho_suboptimal}
+	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-2s/(2(s - 1) + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{p(d)}} \biggr) \vee \frac{1}{n},
+	\end{equation}
+	then $\Ebb_{f_0}[1 - \phi] \leq b$.
+\end{theorem}
+We do not know whether the culprit here is the test itself, our some looseness in our theoretical bounds, and leave the matter for future work.
 
@@
 
-\paragraph{Least squares.}
+\paragraph{Nonparametric least squares.}
 The standard recommended alternative to spectral series methods, when the distribution $P$ is considered non-uniform or unknown, is to do least-squares.
@@
 As mentioned at the outset, our work can be viewed as a contribution both to the fields of nonparametric regression with series estimators, and to graph-based learning.
-Fitting, we end our discussion by mentioning some open work in each of these directions.
+Fittingly, we end our discussion by mentioning some open work in each of these directions.
 

### 2021-03-19 code/estimation_mse_pipeline.R
Added a new 1d regression function for simulation of Laplacian smoothing.

@@ -7,7 +7,7 @@ source("graph.R")
 source("estimators.R")
 
 # Please change this line to whichever config you wish to run.
-source("configs/estimation_mse_config3.R")
+source("configs/estimation_wigglycosine1d.R")
 
 #----------------------------------------------------#
 # This is the pipeline for running regression estimation experiments.
@@ -18,6 +18,7 @@ best_fits_by_method <- vector(mode = "list",length = length(ns))
 mse <- vector(mode = "list",length = length(ns))
 Xs <- vector(mode = "list", length = length(ns))
 f0s <- vector(mode = "list", length = length(ns))
+Ys <- vector(mode = "list", length = length(ns))
 thetas <- vector(mode = "list", length = length(ns))
 
 for(ii in 1:length(ns))
@@ -84,6 +85,7 @@ for(ii in 1:length(ns))
   # Taking the last iteration X, it doesn't matter. 
   Xs[[ii]] <- X
   f0s[[ii]] <- f0_evaluations
+  Ys[[ii]] <- Y
 }
 
 ### Save. ###
@@ -102,4 +104,5 @@ save(best_fits_by_method, file = paste0(save_directory, '/best_fits_by_method.R'
 save(thetas, file = paste0(save_directory, '/thetas.R'))
 save(mse, file = paste0(save_directory, '/mse.R'))
 save(Xs, file = paste0(save_directory, '/Xs.R'))
+save(Ys, file = paste0(save_directory, '/Ys.R'))
 save(f0s, file = paste0(save_directory, '/f0s.R'))

### 2021-03-19 code/plot.R

@@ -3,7 +3,7 @@ library(ggplot2)
 library(gridExtra)
 
 ### Plot. ###
-save_directory <- "data/AISTATS/cosine/d5_largesample" # Please change this to whichever directory you prefer.
+save_directory <- "data/20210318210219" # Please change this to whichever directory you prefer.
 plot_directory <- file.path(save_directory,"plots")
 dir.create(plot_directory)
 
@@ -15,6 +15,7 @@ load(file.path(save_directory,"thetas.R"))
 load(file.path(save_directory,"mse.R"))
 load(file.path(save_directory,"Xs.R"))
 load(file.path(save_directory,"f0s.R"))
+load(file.path(save_directory,"Ys.R"))
 
 # Plot of mse by tuning parameter.
 pdf(file.path(plot_directory,"mse.pdf"))
@@ -55,30 +56,54 @@ for(jj in 1:length(methods))
 }
 dev.off()
 
+# Plot of true regression function
+
+# Parameters
+xlim = c(-1,1)
+ylim = c(min(unlist(Ys)),max(unlist(Ys)))
+col = "grey34"
+cex = .6
+lwd = 1.1
+cex.main = 2.5
+cex.lab = 2.5
+cex.axis = 2
+
+if(d == 1)
+{
+  for(ii in 1:length(ns))
+  {
+    pdf(file.path(plot_directory,paste0("regression_function_",ii,".pdf")))
+    plot_df <- data.frame(x = Xs[[ii]],f0 = f0s[[ii]], y = Ys[[ii]])
+    plot(x = plot_df$x,y = plot_df$y, 
+         col = col, 
+         cex = cex, cex.main = cex.main, cex.lab = cex.lab, cex.axis = cex.axis,
+         xlim = xlim, ylim = ylim, lwd = lwd,
+         xlab = "",ylab = "", main = "True function")
+    lines(x = plot_df$x[order(plot_df$x)],y = plot_df$f0[order(plot_df$x)],lwd = 1.5)
+    dev.off()
+  }
+}
+
 # Plot of fitted values for best choice of tuning parameter.
 if(d == 1)
 {
-  pdf(file.path(plot_directory,"estimates.pdf"))
   for(jj in 1:length(methods))
   {
-    fitted_value_plots <- vector(mode = "list",length = length(ns))
     for(ii in 1:length(ns))
     {
-      plot_df <- data.frame(x = Xs[[ii]],f = f0s[[ii]], fhat = best_fits_by_method[[ii]][[jj]])
[…]

### 2021-03-19 code/sample.R

@@ -1,8 +1,27 @@
+make_sample_gaussian <- function(d)
+{
+  sample_X <- function(n){
+    X <- matrix(NA,nrow = n,ncol = d)
+    while(any(is.na(X))){
+      X[which(is.na(X))] <- rnorm(sum(is.na(X)))
+      X[abs(X) > 1] <- NA
+    }
+    return(X)
+  }
+}
+
 make_sample_uniform <- function(d)
 {
   sample_X <- function(n){
     X <- matrix(runif(n*d,-1,1),ncol = d)
     return(X)
+  }
+}
+
+make_wiggly_gaussian <- function(d,n)
+{
+  f0 <- function(x){
+    return(prod((1/dnorm(x)) * cos(2*pi*x/dnorm(x)^{1.5})))
   }
 }
 

### 2021-03-19 talks/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
AISTATS oral presentation slides nearly complete.

@@
-\documentclass{beamer}
+\documentclass[xcolor=dvipsnames]{beamer}
 
@@
 \setbeamertemplate{itemize items}[circle]
+
+\definecolor{Red}{rgb}{.9,.05,.05}
+\setbeamercolor{alerted text}{fg=Red}
 \usefonttheme[onlymath]{serif}
@@
 \begin{frame}{Basic problem setup}
-\alert{Nonparametric regression, random design}: We observe $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are sampled independently from $P$, and
+\alert{Nonparametric regression, random design}: We observe $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are sampled independently from density $p$, and
 \begin{equation*}
@@
 \begin{frame}{Laplacian smoothing}
-Laplacian smoothing solves a discrete, penalized least squares problem. 
+Laplacian smoothing solves a discrete, penalized least squares problem on a graph.
 \begin{enumerate}
-	\item  Form a \alert{neighborhood graph} $G$ over $X_1,\ldots,X_n$, with weighted edges $W_{i,j} = K(\|X_i - X_j\|/r)$. The \alert{graph Laplacian} $L$ is defined by
+	\item  Form a \alert{neighborhood graph} $G$ over $X_1,\ldots,X_n$, with weighted edges $W_{ij} = K(\|X_i - X_j\|/r)$. The \alert{graph Laplacian} $L$ is defined by
 	\begin{equation*}
-	(Lf)_i = \sum_{i = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr) W_{i,j}.
+	(Lf)_i = \sum_{i = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr) W_{ij}.
 	\end{equation*}
@@
 	\wh{f} = \argmin_{f \in \Reals^n} \|Y - f\|_n^2 + \frac{\lambda}{2} f^{\top} L f, \end{equation*}
-	with corresponding test statistic $\wh{T} = \|\wh{f}\|_n^2$.
+	%with corresponding test statistic $\wh{T} = \|\wh{f}\|_n^2$.
 \end{enumerate}
-\textcolor{red}{(TODO): Simple picture here.}
+\begin{figure}
+	\centering
+	\includegraphics[width=.25\textwidth]{figures/wiggly_cosine/regression_function_1.pdf}
+	\includegraphics[width=.25\textwidth]{figures/wiggly_cosine/laplacian_smoothing_estimate_1.pdf} 
+\end{figure}
 \end{frame}
@@
 \begin{equation*}
-f^T L f = \frac{1}{2}\sum_{i,j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 W_{i,j} \approx n^2 r^{d + 2} \int \|\nabla f\|_2^2 \,dx.
+f^T L f = \frac{1}{2}\sum_{i,j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 W_{i,j} \approx n^2 r^{d + 2} \int \|\nabla f\|_2^2 p^2(x) \,dx.
 \end{equation*}
-Seemingly $\wh{f}$ is a noisy approximation of a very classical estimator $\wt{f}$: aka \alert{smoothing spline} (1d), or \alert{thin-plate spline} ($d > 1$), defined as
+So $\wh{f}$ seems like a noisy approximation of a very classical estimator $\wt{f}$: aka \alert{smoothing spline} ($d = 1$), or \alert{thin-plate spline} ($d \geq 2$), defined as
 \begin{equation*}
@@
 \begin{frame}{Main Question}
-In $1d$, smoothing splines achieve the minimax optimal rate \blue{$n^{-2/3}$} (estimation) or \blue{$n^{-4/5}$} (testing).
+In $1d$, smoothing splines achieve the minimax optimal rate \blue{$n^{-2/3}$} (estimation) or \blue{$n^{-4/5}$} (testing) over Sobolev functions.
 \begin{block}{{\bf Question}}
@@
 		\item Laplacian smoothing has error of \blue{$n^{-2/3}$} (estimation) and \blue{$n^{-4/5}$} (testing) when $d = 1$, matching smoothing splines.
-		\item Laplacian smoothing also achieves optimal error \blue{$n^{-2/(2 + d)}$} (estimation) and \blue{$n^{-4/(4 + d)}$} (testing) when $d = 2,3,4$ (up to $\log(n)$ factors when $d = 4$).
-		\item On the other hand $\wt{f}$ is \alert{not even well-defined} when $d > 1$. Well known in spline theory community.
+		\item Laplacian smoothing also achieves optimal error \blue{$n^{-2/(2 + d)}$} (estimation) and \blue{$n^{-4/(4 + d)}$} (testing) when $d = 2,3,4$ (up to $\log n$ factor when $d = 4$).
+		\item On the other hand $\wt{f}$ is \alert{not even well-defined} when $d > 1$. Well known in spline/semi-supervised learning communities.
 	\end{itemize}
@@
 		\begin{tabular}{p{.25\textwidth} | p{.25\textwidth} p{.25\textwidth} }
-			Dimension & Laplacian smoothing & Thin-plate splines \\
+			Dimension & Laplacian smoothing & Smoothing/Thin-plate splines \\
 			\hline
-			$d = 1$ & \blue{${n^{-2/3}}$} & \blue{${n^{-2/3}}$} \\
-			$d = 2,3$ & \blue{${n^{-2/(2 + d)}}$} & \red{NA} \\
-			$d = 4$ & $\blue{{n^{-1/3}}} (\log n)^{1/3}$ & \red{NA} \\
-			$d \geq 5$  & $(\log n/n)^{4/(3d)}$ & \red{NA} \\
+			$d = 1$ & \blue{\bm{${n^{-2/3}}$}} & \blue{${n^{-2/3}}$} \\
+			$d = 2,3$ & \blue{\bm{${n^{-2/(2 + d)}}$}} & \red{NA}/\red{$1$} \\
+			$d = 4$ & $\bm{\blue{{n^{-1/3}}} (\log n)^{1/3}}$ & \red{NA}/\red{$1$} \\
+			$d \geq 5$  & $\bm{(\log n/n)^{4/(3d)}}$ & \red{NA}/\red{$1$} \\
 		\end{tabular}
@@
 		\begin{tabular}{p{.25\textwidth} | p{.25\textwidth} p{.25\textwidth} }
-			Dimension & Laplacian smoothing & Thin-plate splines \\
+			Dimension & Laplacian smoothing & Smoothing/Thin-plate splines \\
 			\hline
-			$d = 1$ & \blue{${n^{-4/5}}$} & \blue{${n^{-4/5}}$} \\
-			$d = 2,3$ & \blue{${n^{-4/(4 + d)}}$} & \red{NA} \\
-			$d \geq 4$ & $\blue{{n^{-1/2}}}$ & \red{NA} \\
+			$d = 1$ & \blue{\bm{${n^{-4/5}}$}} & \blue{${n^{-4/5}}$} \\
+			$d = 2,3$ & \blue{\bm{${n^{-4/(4 + d)}}$}} & \red{NA}/\red{$n^{-1/2}$} \\
+			$d \geq 4$ & $\blue{\bm{{n^{-1/2}}}}$ & \red{NA}/\blue{$n^{-1/2}$} \\
 		\end{tabular}
@@
 \end{itemize}
-As a result the criterion underlying $\wt{f}$ is \red{not continuous} with respect to $H^1$ norm. I can construct sequences of ``bump functions'' $f_1,f_2,\ldots$ such that
+As a result the criterion underlying $\wt{f}$ is \alert{not continuous} with respect to $H^1$ norm. \blue{(Green and Silverman)} construct sequences of ``bump functions'' $f_1,f_2,\ldots$ such that
 \begin{equation*}
-f_k(X_i) = Y_i \quad \text{for all $k$, but} \quad \|f_k\|_{H^1} \to 0.
+f_k(X_i) = Y_i \quad \text{for all $i = 1,\ldots,n$, but} \quad \|f_k\|_{H^1} \to 0.
 \end{equation*}
-Implication: For any $\delta > 0$, there exists a function $f$ which satisfies
-\begin{equation*}
-f(X_i) = Y_i \quad \text{for all $i = 1,\ldots,n$, yet} \int\|\nabla f(x)\|^2 \,dx < \delta. 
-\end{equation*} 
+\alert{Bottom line}: The continuum thin-plate spline problem assigns (much) higher loss + penalty to the regression function $f_0$ than to bump functions $f_k$.
 \end{frame}
@@
 \begin{frame}{Explanation: Success of Laplacian Smoothing}
-Roughly, discretization is a blessing.
-
+\alert{Discretization} is a blessing!  
+\begin{itemize}
+	\item Previous construction of ``bump functions'' $f_1,f_2,\ldots$ will satisfy
+	\begin{equation*}
+	f_k^{\top} L f_k = \sum_{i,j = 1}^{n} (Y_i - Y_j)^2 K(\|X_i - X_j\|/r) \asymp n^2r^{d}.
+	\end{equation*}
+	\item If $f_0$ is Sobolev smooth, then $f_0^{\top} L f_0 \lesssim n^2r^{d + 2}$.
+	\item Thus for $\lambda \gg (nr^{d + 2})^{-1} r^{-2}$, we have
+	\begin{equation*}
+	\|Y - f_k\|_n^2 + \lambda f_k^{\top} L f_k \gg \|Y - f_0\|_n^2 + \lambda f_0^{\top} L f_0.
+	\end{equation*}
+\end{itemize}
+\alert{Bottom line}: the discrete problem assigns (much) higher loss + penalty to bump functions $f_k$ than to regression function $f_0$.
 \end{frame}
@@
 \begin{frame}{What about $d > 4$?}
-content...
+\alert{Estimation}: the class $H^1(G;M) = \{f: f^{\top}Lf \leq M\}$ is too large to show $\wh{f}$ is optimal.
+Similar phenomena in ERM over H\"{o}lder classes when $d > 2$ \blue{(Birge and Massart)}, and fixed (grid) graphs \blue{(Sadhanala et al.)}.
+Empirically, we see that for $f_0$ simple $\cos$ function,
+\begin{figure}[tb]
+	\includegraphics[width=.24\textwidth]{figures/cosine/mse_by_sample_size_2d.pdf}
+	\includegraphics[width=.24\textwidth]{figures/cosine/mse_by_sample_size_3d.pdf} 
+	\includegraphics[width=.24\textwidth]{figures/cosine/mse_by_sample_size_4d.pdf}
+	\includegraphics[width=.24\textwidth]{figures/cosine/mse_by_sample_size_5d.pdf}
+\end{figure}
+
+\alert{Testing}: Sobolev class $H^1$ does not embed into $L^4$ (Sobolev Embedding Theorem). Usual test statistics may not have a variance.
 \end{frame}
@@
 \begin{frame}{Extensions}
+Several different directions.
+\begin{itemize}
+	\item \alert{Manifold adaptivity}: When $p$ has support on a manifold of dimension $m < d$, optimal rates are \blue{$n^{-2/(2 + m)}$} (estimation) and \blue{$n^{-4/(4 + m)}$} (testing).
+Does Laplacian smoothing automatically adapt?
+\blue{(Yes!
+See paper.)}
+	\item \alert{Out-of-sample estimation}: Can we extend estimator $\wh{f}$ to have optimal out-of-sample mean squared error? \blue{(Yes, if $f_0$ is Lipschitz! See paper.)}
+	\item \alert{Smoothness adaptivity}: When $f_0$ has $s$ smooth derivatives, optimal rates are \blue{$n^{-2s/(2s + m)}$} (estimation) and \blue{$n^{-4s/(4s + m)}$} (testing).
+Does Laplacian smoothing adapt?
+(Future work...)
+	\item \alert{Different methods}: Spectral cutoff? $k$NN graph? Normalized Laplacian? (Future work...)
+\end{itemize}
+\end{frame}
+
+\begin{frame}[t]{Summary}
+	Specific takeaway: Laplacian smoothing is an optimal method for nonparametric regression. \newline
+
+	General takeaway: Discretization can allow us to solve problems even when their continuum analogues are not well-posed.
+	\newline 
+	\newline
+	\newline
+
+	\begin{center}
+	{\huge Thank You!}
+	\end{center}
 

### 2021-03-19 talks/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
Final version of AISTATS oral presentation.

@@
 \usepackage{bm}
+\usepackage{wasysym}
 
@@
 \usetheme{default}
+\setbeamerfont{author}{size = \large}
+\setbeamerfont{institute}{size=\large}
+\setbeamerfont{footline}{series = \bfseries}
 \setbeamertemplate{navigation symbols}{}
@@
 \setbeamertemplate{itemize items}[circle]
-
 \definecolor{Red}{rgb}{.9,.05,.05}
@@
 
-\title{Laplacian Smoothing on Neighborhood Graphs}
+\title{Minimax Optimality of Laplacian Smoothing}
+\subtitle{\small (Full title: Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs)}
 
@@
 \institute{Department of Statistics and Data Science \\
-		   Carnegie Mellon Unviersity}
+\includegraphics[width=.5 \textwidth]{figures/cmu.png}}
 \date{}
@@
 \titlepage
+\centering \emph{Co-authors: Sivaraman Balakrishnan, Ryan Tibshirani}
 \end{frame}
+
+\addtobeamertemplate{navigation symbols}{}{
+	\usebeamerfont{footline}
+	\usebeamercolor[fg]{footline}
+	\hspace{1em}
+	\insertframenumber
+}
+\addtocounter{framenumber}{-1}
 
@@
 \end{equation*}
-Either \alert{estimate} $f_0$ with some $\hat{f}$, or \alert{test} whether $f_0 = 0$.
+Either \alert{estimate} $f_0$ with some $\wh{f}$, or \alert{test} whether $f_0 = 0$.
 \end{frame}
@@
 	\begin{equation*}
-	\wh{f} = \argmin_{f \in \Reals^n} \|Y - f\|_n^2 + \frac{\lambda}{2} f^{\top} L f, \end{equation*}
+	\wh{f} = \argmin_{f \in \Reals^n} \|Y - f\|_n^2 + \frac{\rho}{2} f^{\top} L f,\quad\textrm{with test statistic}\quad\wh{T} = \|\wh{f}\|_n^2. 
+	\end{equation*}
 	%with corresponding test statistic $\wh{T} = \|\wh{f}\|_n^2$.
@@
 \end{itemize}
-Intuition: when $n$ large, $r$ small and $f$ smooth, \textcolor{blue}{(Bousquet et al.)} show
+Intuition: when $n$ large, $r$ small and $f$ smooth, {(Bousquet et al. 2004)} show
 \begin{equation*}
@@
 \end{equation*}
-So $\wh{f}$ seems like a noisy approximation of a very classical estimator $\wt{f}$: aka \alert{smoothing spline} ($d = 1$), or \alert{thin-plate spline} ($d \geq 2$), defined as
+So $\wh{f}$ seems like a noisy and discrete approximation to a very classical estimator $\wt{f}$: aka \alert{smoothing spline} ($d = 1$), or \alert{thin-plate spline} ($d \geq 2$), defined as
 \begin{equation*}
-\widetilde{f} = \argmin_{f \in H^1} \|Y - f\|_n^2 + \lambda \int \|\nabla f(x)\|_2^2  \,dx.
+\widetilde{f} = \argmin_{f \in H^1} \|Y - f\|_n^2 + \rho \int \|\nabla f(x)\|_2^2  \,dx.
 \end{equation*}
@@
 \end{itemize}
-As a result the criterion underlying $\wt{f}$ is \alert{not continuous} with respect to $H^1$ norm. \blue{(Green and Silverman)} construct sequences of ``bump functions'' $f_1,f_2,\ldots$ such that
+As a result the criterion underlying $\wt{f}$ is \alert{not continuous} with respect to $H^1$ norm. (Green and Silverman 1994) construct sequences of ``bump functions'' $f_1,f_2,\ldots$ such that
 \begin{equation*}
@@
 \end{equation*}
-\alert{Bottom line}: The continuum thin-plate spline problem assigns (much) higher loss + penalty to the regression function $f_0$ than to bump functions $f_k$.
+\alert{Bottom line}: The continuum thin-plate spline problem assigns (much) higher loss + penalty to the regression function $f_0$ than to bump functions $f_k$. \red{\frownie} 
 \end{frame}
@@
 	\begin{equation*}
-	\|Y - f_k\|_n^2 + \lambda f_k^{\top} L f_k \gg \|Y - f_0\|_n^2 + \lambda f_0^{\top} L f_0.
+	\|Y - f_k\|_n^2 + \rho f_k^{\top} L f_k \gg \|Y - f_0\|_n^2 + \rho f_0^{\top} L f_0.
 	\end{equation*}
 \end{itemize}
-\alert{Bottom line}: the discrete problem assigns (much) higher loss + penalty to bump functions $f_k$ than to regression function $f_0$.
+\alert{Bottom line}: the discrete problem assigns (much) higher loss + penalty to bump functions $f_k$ than to regression function $f_0$. \blue{\smiley}
 \end{frame}
@@
 \alert{Estimation}: the class $H^1(G;M) = \{f: f^{\top}Lf \leq M\}$ is too large to show $\wh{f}$ is optimal.
-Similar phenomena in ERM over H\"{o}lder classes when $d > 2$ \blue{(Birge and Massart)}, and fixed (grid) graphs \blue{(Sadhanala et al.)}.
+Similar phenomena in ERM over H\"{o}lder classes when $d > 2$ {(Birge and Massart 1993)}, and fixed (grid) graphs {(Sadhanala et al.
+2016)}.
 Empirically, we see that for $f_0$ simple $\cos$ function,
 \begin{figure}[tb]
-	\includegraphics[width=.24\textwidth]{figures/cosine/mse_by_sample_size_2d.pdf}
-	\includegraphics[width=.24\textwidth]{figures/cosine/mse_by_sample_size_3d.pdf} 
-	\includegraphics[width=.24\textwidth]{figures/cosine/mse_by_sample_size_4d.pdf}
-	\includegraphics[width=.24\textwidth]{figures/cosine/mse_by_sample_size_5d.pdf}
+	\includegraphics[width=.24\textwidth]{figures/cosine/mse_by_sample_size_2d.png}
+	\includegraphics[width=.24\textwidth]{figures/cosine/mse_by_sample_size_3d.png} 
+	\includegraphics[width=.24\textwidth]{figures/cosine/mse_by_sample_size_4d.png}
+	\includegraphics[width=.24\textwidth]{figures/cosine/mse_by_sample_size_5d.png}
 \end{figure}
@@
 \begin{frame}{Analysis}
+We use a conditional-on-design \alert{bias-variance} decomposition, based on the eigenvalues $\lambda_k$ of $L$. \newline
+
+For estimation,
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 \lesssim \frac{f_0^{\top} L f_0}{n\lambda_k} + \frac{1}{n} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2}.
+\end{equation*}
+
+For testing, level-$\alpha$ test $\varphi = 1\{\wh{T} \geq t_{\alpha}\}$ has negligible Type II error when
+\begin{equation*}
+\|f_0\|_n^2 \gtrsim \frac{f_0^{\top} L f_0}{n\lambda_k} + \frac{1}{n} \sum_{k = 1}^{n} \sqrt{\frac{1}{(\rho \lambda_k + 1)^4}}.
+\end{equation*}
+
+Then we make use of some recent results (Burago et al. 2014, Garcia Trillos et al. 2019) on convergence of eigenvalues $\lambda_k$, and derive some of our own.
 \end{frame}

### 2021-03-19 code/plot.R
Trivial change to plotting code.

@@ -3,7 +3,7 @@ library(ggplot2)
 library(gridExtra)
 
 ### Plot. ###
-save_directory <- "data/20210318210219" # Please change this to whichever directory you prefer.
+save_directory <- "data/AISTATS/cosine/d5_largesample" # Please change this to whichever directory you prefer.
 plot_directory <- file.path(save_directory,"plots")
 dir.create(plot_directory)
 
@@ -15,7 +15,7 @@ load(file.path(save_directory,"thetas.R"))
 load(file.path(save_directory,"mse.R"))
 load(file.path(save_directory,"Xs.R"))
 load(file.path(save_directory,"f0s.R"))
-load(file.path(save_directory,"Ys.R"))
+# load(file.path(save_directory,"Ys.R"))
 
 # Plot of mse by tuning parameter.
 pdf(file.path(plot_directory,"mse.pdf"))
@@ -72,7 +72,7 @@ if(d == 1)
 {
   for(ii in 1:length(ns))
   {
-    pdf(file.path(plot_directory,paste0("regression_function_",ii,".pdf")))
+    png(file.path(plot_directory,paste0("regression_function_",ii,".png")))
     plot_df <- data.frame(x = Xs[[ii]],f0 = f0s[[ii]], y = Ys[[ii]])
     plot(x = plot_df$x,y = plot_df$y, 
          col = col, 
@@ -91,8 +91,8 @@ if(d == 1)
   {
     for(ii in 1:length(ns))
     {
-      plot_name <- paste0(names(methods)[jj],"_estimate_",ii,".pdf")
-      pdf(file.path(plot_directory,plot_name))
+      plot_name <- paste0(names(methods)[jj],"_estimate_",ii,".png")
+      png(file.path(plot_directory,plot_name))
       plot_df <- data.frame(x = Xs[[ii]], y = Ys[[ii]],fhat = best_fits_by_method[[ii]][[jj]])
       plot(x = plot_df$x,y = plot_df$y, 
            col = col,
@@ -107,8 +107,8 @@ if(d == 1)
 }
 
 # Plot of mse---for best choice of tuning parameter---by sample size
-pdf(file.path(plot_directory,"mse_by_sample_size.pdf"))
-par(mar = c(5.1,4.5,4.1,4.1)) # The left hand side was getting cut off a tad.
+png(file.path(plot_directory,"mse_by_sample_size.png"))
[…]

### 2021-03-21 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Finished editing the section on rates for manifold domain.

@@
 			$s \leq 3$ & $\bm{n^{-2s/(2s + d)}}$ & $\bm{n^{-2s/(2s + m)}}$ \\
-			$s > 3$  & $\bm{n^{-2s/(2s + d)}}$ & \textcolor{red}{(?)}
+			$s > 3$  & $\bm{n^{-2s/(2s + d)}}$ & $n^{-6/(6 + m)}$
 		\end{tabular}
@@
 			\hline
-			\multirow{3}{*}{$s > 3$} & $\dim(\mc{X}) \leq 4$ & $\bm{n^{-4s/(4s + d)}}$ & \textcolor{red}{(?)} \\
-			& $4 < \dim(\mc{X}) < 4s$ & \textcolor{red}{(?)} & \textcolor{red}{(?)} \\
+			\multirow{3}{*}{$s > 3$} & $\dim(\mc{X}) \leq 4$ & $\bm{n^{-4s/(4s + d)}}$ & $n^{-12/(12 + d)}$ \\
+			& $4 < \dim(\mc{X}) < 4s$ & $n^{-2s/(2(s - 1) + d)}$ & $n^{-6/(4 + m)}$ \\
 			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
@@
 When $4s \geq d$ the functions in $H^s(\Xset)$ are very irregular---formally speaking $H^s(\Xset)$ does not continuously embed into $\Leb^4(\Xset)$ when $4s \geq d$---and as far as we know, the minimax testing rates in this regime have not been worked out.
-However, if one explicitly assumes that $f_0 \in \Leb^4(\mc{X})$ then the critical radius is given by $\epsilon^2 \asymp Mn^{-1/2}$; as a sanity check, note that this is strictly worse than the rate in~\eqref{eqn:sobolev_space_testing_critical_radius}.
+However, if one explicitly assumes that $f_0 \in \Leb^4(\mc{X})$ then the critical radius is given by $\epsilon^2 \asymp Mn^{-1/2}$.
+As a sanity check note that this is strictly worse than the rate in~\eqref{eqn:sobolev_space_testing_critical_radius}.
 As we discuss after our first main theorem regarding testing (Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}), this rate is achievable by a test based on $\wh{T}$.
@@
 As previously explained, Laplacian eigenmaps is a discrete and noisy approximation to a spectral series method, of particular use when the design distribution $P$ is unknown.
-The hope is that the error incurred by this approximation is sufficiently small such that Laplacian eigenmaps preserves the optimality properties of the spectral series methods they are approximating.
-In this section and the following one, we will show that this is indeed the case; the estimator $\wh{f}$, and a test using the statistic $\wh{T}$, achieve optimal estimation and goodness-of-fit testing rates over Sobolev classes.
+Our goal is to show that Laplacian eigenmaps are rate optimal methods, notwithstanding the extra error incurred by this approximation.
+In this section and the following one, we will see that this is indeed the case: the estimator $\wh{f}$, and a test using the statistic $\wh{T}$, achieve optimal estimation and goodness-of-fit testing rates over Sobolev classes.
 
@@
 However, because the estimator $\wh{f}$ is defined only in-sample, we cannot use such methods to select the graph radius $\varepsilon$, or number of eigenvectors $K$.
-We return to this issue upon considering out-of-sample extensions of $\wh{f}$ later, in Section~\ref{sec:out_of_sample}.
+We return to this issue in Section~\ref{sec:out_of_sample}, when we propose an out-of-sample extension of $\wh{f}$.
 	\item The upper bound given in Equation~\eqref{eqn:laplacian_eigenmaps_estimation_fo} holds with probability $1 - \delta - Cn\exp(-cn\varepsilon^d)$.
@@
 As $s \to \infty$ and $d$ remains fixed, the optimal estimation and testing rates approach the parametric rate $n^{-1}$.
-On the other hand, the sharpest known results show that Laplacian eigenvectors $v_k$ converge to continuum eigenfunctions $\psi_k$ only at a rate of \textcolor{red}{(?)}, which is obviously much slower than $n^{-1}$.
+On the other hand, the sharpest known results show that Laplacian eigenvectors $v_k$ converge to continuum eigenfunctions $\psi_k$---and thus, for a fixed value of $k$, $\wh{f}$ converges to $\wt{f}$---only at a rate of \textcolor{red}{(?)}, which is obviously much slower than $n^{-1}$.
 
@@
 Consequently, exactly computing the eigenvectors $v_k$ is more computationally intensive than if the graph radius were on the order of the connectivity threshold $\varepsilon \asymp (\log n/n)^{1/d}$.
-Recall that in the first-order case, we established in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that this smaller choice of $\varepsilon$ still resulted in an optimal estimator.
+Recall that in the first-order case, we established in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that as long as $\varepsilon \gg (\log n/n)^{1/d}$---meaning average degree growing like $\omega(\log n)$---the estimator $\wh{f}$ could still be optimal.
 Thus Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and~\ref{thm:laplacian_eigenmaps_estimation_ho} can be viewed as revealing a tradeoff between statistical and computational efficiency; although to be clear, we have no theoretical evidence that Laplacian eigenmaps \emph{fails} to adapt to higher-order smoothness when $\varepsilon \asymp (\log n/n)^{1/d}$---we simply cannot prove that it succeeds.
@@
 	The graph $\wc{G}$ is said to be $\sigma$-spectrally similar to $G$.
-Let $\wc{f}$ be the Laplacian eigenmaps estimator, computed with respect to the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ .
-Because $\wc{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wc{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$ \textcolor{red}{(reference needed)}.
+Let $\wc{f}$ be the Laplacian eigenmaps estimator computed using the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ .
+Because $\wc{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wc{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$, and consequently much faster to compute $\wc{f}$ than $\wh{f}$\textcolor{red}{(reference needed)}.
 Statistically speaking, as we show in Appendix~\textcolor{red}{(?)} $\wc{f}$ has $L^2(P_n)$-error of at most $\sigma^{2s}$ times the error of $\wh{f}$.
 Therefore, as long as $\sigma \asymp 1$ the estimator $\wc{f}$ will also be rate-optimal.
-Indeed, as pointed out by~\citet{sadhanala16b}, geometric graphs such as $G$ exhibit a special structure.
-In this case there exist far simpler and faster methods for sparsification than those of~\textcolor{red}{(references)}, which, at least empirically, appear to suffice.
-	\item The requirement $p \in C^{s - 1}(\mc{X})$ is a strong condition, but is essential to showing that $\wh{f}$ enjoys faster rates of convergence.
+Finally, geometric graphs such as $G$ exhibit a special structure; very roughly speaking, no one edge is a bottleneck in geometric graphs.
+As pointed out by~\citet{sadhanala16b}, in this case there exist far simpler and faster methods for sparsification than those of~\textcolor{red}{(references)}; at least empirically, these methods appear to do the job.
+	\item The requirement $p \in C^{s - 1}(\mc{X})$ is a strong condition, but is essential to showing that $\wh{f}$ enjoys faster rates of convergence when $s > 1$.
 We give more detail about this when we discuss our analysis in Section~\ref{subsec:analysis}.
@@
 \end{enumerate}
-When $d \leq 4$, for any value of $s$ it is possible to choose $\varepsilon$ and $K$ such that both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho} are satisfied, and our next theorem establishes that $\varphi$ is an optimal test.
+When $d \leq 4$, for any value of $s \in \mathbb{N}$ it is possible to choose $\varepsilon$ and $K$ such that both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho} are satisfied, and our next theorem establishes that in this situation $\varphi$ is an optimal test.
 In what follows, let $p(d) = \min\{1,2s/d\}$.
@@
 Then $\mathbb{E}_0[\varphi] \leq a$, i.e $\varphi$ is a level-$a$ test.
-Suppose additionally $f_0 \in H_0^s(P,M)$,that $p \in C^{s-1}(\mc{X})$, and that $d \leq 4$.
+Suppose additionally $f_0 \in H_0^s(P,M)$, that $p \in C^{s-1}(\mc{X})$, and that $d \leq 4$.
 Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_testing_ho}, and if $f_0$ satisfies
@@
 \end{theorem}
-Just as in the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that $\varphi$ is a minimax optimal test over $H_0^s(P)$ when $n^{-1/2} \ll M^2 \ll n^{1/4}$ (if $2s < d$), or if $n^{-1/2} \ll M^2 \ll (2/d - 1/2)$ (if $2s \geq d$.)
-
-However, unlike the first-order case, when $4 < d < 4s$ the optimal testing rate over $H_0^s(P)$ continues to be $n^{-4s/(4s + d)}$.
-Unfortunately, we cannot claim that $\varphi$ is an optimal test in this regime.
-The culprit is that there no longer exist any choices of $\varepsilon,K$ which satisfy both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho}, and we can therefore no longer balance testing bias and variance (defined momentarily in~\textcolor{red}{(?)}).
-From Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, we can gather that $\varphi$ must have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2s + d)}$.
+Similarly to the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that $\varphi$ is a minimax optimal test over $H_0^s(P)$ when $n^{-1/2} \ll M^2 \ll n^{1/4}$ (if $2s < d$), or if $n^{-1/2} \ll M^2 \ll (2/d - 1/2)$ (if $2s \geq d$.)
+
+However, unlike the first-order case, when $4 < d < 4s$ the minimax testing rate over $H_0^s(P)$ is still $n^{-4s/(4s + d)}$.
+Unfortunately, we can no longer claim that $\varphi$ is an optimal test in this regime.
+There do not exist any choices of $\varepsilon$ and $K$ which satisfy both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho}.
+As a result we can't balance testing bias and variance (defined momentarily in~\eqref{eqn:testing_biasvariance}).
+It follows from Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} and the triangle inequality that, choosing $K \asymp (n)^{-2s/(2s + d)}$ as is optimal for estimation, $\varphi$ must have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2s + d)}$.
 In fact, we can derive slightly better (though still not optimal) rates.
@@
 \end{theorem}
-We do not know whether the culprit here is the test itself, our some looseness in our theoretical bounds, and leave the matter for future work.
+We do not know whether the problem here is the test itself, our some looseness in our theoretical bounds, and leave the matter for future work.
+
+A last point: as mentioned at the beginning of this section, the estimation and testing rates we obtain in Theorems~\ref{thm:laplacian_eigenmaps_estimation_ho}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} are superior to those implied by the concentration of graph Laplacian eigenvectors $v_k$ around eigenfunctions $\psi_k$ of the weighted Laplace-Beltrami operator $\Delta_P$.
+This implies (at least) one out of the two following explanations.
+\begin{enumerate}
+	\item The convergence rates of $v_k \to \psi_k$ are faster than currently known,  particularly when $p$ exhibits higher-order smoothness.
+	\item The approximation properties of $v_1,\ldots,v_K$ are much stronger than is implied by the concentration of each individual eigenvector $v_k$ around its continuum counterpart $\psi_k$.
+This is reminiscent of similar phenomena exhibited by kernel machines \textcolor{red}{(Belkin 2019)}.
+\end{enumerate}
+Identifying which explanation is responsible would, in our view, be a useful step towards understanding the success of graph Laplacian methods.
+\textcolor{red}{(TODO): Too vague and speculative.
+Maybe worth deleting entirely?}
 
@@
 
-\textbf{(1) Bias and variance terms, for estimation and testing}.
-We analyze the estimation error of $\wh{f}$, and the testing error of $\wh{\varphi}$, by first conditioning on the design points $X$ and deriving \emph{design-dependent} bias and variance terms.
-For estimation, we have that with high probability, conditional on the design $X$,
-\begin{equation*}
-\|\wh{f} - f_0\|_n^2 \leq C\frac{\dotp{L^s f_0}{f_0}_n}{\lambda_{\kappa}^s} + \frac{K}{n}
-\end{equation*}
-For testing, we have~\textcolor{red}{(...)}.
-
-The semi-norm $\dotp{L^s f_0}{f_0}_n$ and eigenvalue $\lambda_{K}$ are random variables that depend the random design points $X_1,\ldots,X_n$.
-It remains to establish suitable upper and lower bounds on these quantities.
-
-\textbf{(2) Upper bound on the semi-norm. } Proposition~\ref{prop:graph_seminorm} gives a sufficient upper bound on $\dotp{L^s f_0}{f_0}_n$. Take $\beta := s - 1$. 
+We now outline the high-level strategy we follow when proving each of Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}.
+We analyze the estimation error of $\wh{f}$, and the testing error of $\wh{\varphi}$, by first conditioning on the design points ${\bf X}$ and deriving \emph{design-dependent} bias and variance terms.
+For estimation, we have that with probability at least $1 - \exp(-K)$,
+\begin{equation}
+\label{eqn:estimation_biasvariance}
+\|\wh{f} - f_0\|_n^2 \leq \underbrace{\frac{\dotp{L^s f_0}{f_0}_n}{\lambda_{K}^s}}_{\textrm{bias}} + \underbrace{\frac{5K}{n} \vphantom{\frac{\dotp{L^s f_0}{f_0}_n}{\lambda_{K}^s}}}_{\textrm{variance}}.
+\end{equation}
+For testing, we have that $\varphi$ (which is a level-$a$ test by construction) also has small Type II Error, $\Ebb_{f_0}[1 - \phi] \leq b/2$, if 
+\begin{equation}
+\label{eqn:testing_biasvariance}
+\|f_0\|_n^2 \geq  \underbrace{\frac{\dotp{L^s f_0}{f_0}_n}{\lambda_{K}^s}}_{\textrm{bias}} + \underbrace{32\frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]}_{\textrm{variance}}.
+\end{equation}
+The graph quadratic form $\dotp{L^s f_0}{f_0}_n$, eigenvalue $\lambda_{K}$, and empirical squared norm $\|f_0\|_n^2$ are each random variables that depend the random design points $X_1,\ldots,X_n$.
+We now establish suitable upper and lower bounds on these quantities.
+
+\paragraph{Estimates on graph quadratic forms.}
+We begin by restating in Proposition~\ref{prop:graph_seminorm_fo} an upper bound on the first-order quadratic form $\dotp{L f_0}{f_0}_n$ from \textcolor{blue}{Green et al. 2021}.
+\begin{proposition}[Lemma~1 of \textcolor{blue}{Green et al. 2021}]
+	\label{prop:graph_seminorm_fo}
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H^1(P,M)$.
+There exist constants $c,C$ that do not depend on $f_0$ such that the following statement holds for any $\delta \in (0,1)$: if $\varepsilon < c$, then
+	\begin{equation}
+	\label{eqn:graph_seminorm_fo}
+	\dotp{L_{n,\varepsilon}f_0}{f_0}_n \leq \frac{C}{\delta} M^2,
+	\end{equation}
+	with probability at least $1 - \delta$.
+\end{proposition}
+
+In this work, we establish that a similar result holds for the higher-order case $s > 1$.
+
 \begin{proposition}
-	\label{prop:graph_seminorm} 
-	Suppose $f \in H_0^s(\mc{X})$, that $p \in C^{\beta}(\mc{X})$, and that $r \gtrsim n^{-1/d}$. Then with probability at least $1 - \delta$, it holds that 
-	\begin{equation}
-	\label{eqn:graph_seminorm_1}
-	\dotp{L^s f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^s} \|p\|_{C^{\beta}} + \textcolor{red}{(?)}
-	\end{equation}
-	Therefore, if additionally $r \gtrsim n^{-1/(2\beta + d)}$, then 
-	\begin{equation}
-	\label{eqn:graph_seminorm_2}
-	\dotp{L^s f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^s} \|p\|_{C^{\beta}},
-	\end{equation}
-	with the same probability guarantee. When $s = 1$, the results hold for $f \in H^1(\mc{X})$ and any density $p$ that satisfies Assumption~\ref{asmp:density}.
+	\label{prop:graph_seminorm_ho} 
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f_0 \in H_0^s(P;M)$ and $p \in C^{s - 1}(\mc{X})$.
+Then there exist constants $c$ and $C$ that do not depend on $f_0$ such that the following statement holds for any $\delta \in (0,1)$: if $Cn^{-1/(2(s - 1) + d)} < \varepsilon < c$, then
+	\begin{equation}
+	\label{eqn:graph_seminorm_ho}
+	\dotp{L_{n,\varepsilon}^s f}{f}_n \leq \frac{C}{\delta} M^2 \|p\|_{C^{s}(\mc{X})}^2,
+	\end{equation}
+	with probability at least $1 - \delta$.
 \end{proposition}
-The proof of Proposition~\ref{prop:graph_seminorm} involves two steps.
-In the first step, we show that with probability at least $1 - \delta$, the graph semi-norm $\dotp{L^s f}{f}_n$ is upper bounded by a non-local continuum energy $E_{\varepsilon}^{(s)}(f)$, defined as
-\begin{equation*}
-E_{\varepsilon}^{(s)}(f) := \dotp{L_{P,\varepsilon}^sf}{f}_{P}
-\end{equation*}
-where $L_{P,\varepsilon}$ is a non-local approximation to $\Delta_P$, 
+The proof of Proposition~\ref{prop:graph_seminorm_ho} can be divided into two steps.
+In the first step, we show that when $Cn^{-1/(2(s - 1) + d)}$, then with probability at least $1 - \delta$ the graph semi-norm $\dotp{L^s f}{f}_n$ is upper bounded by $1/\delta$ times a non-local continuum energy $E_{\varepsilon}^{(s)}(f)$, defined as
+\begin{equation*}
+E_{\varepsilon}^{(s)}(f) := \dotp{L_{P,\varepsilon}^sf}{f}_{P}.
+\end{equation*}
+Here $L_{P,\varepsilon}$ is a non-local approximation to $\Delta_P$, 
 \begin{equation}
@@
 \end{equation}
-In the second step, we show that $E_{\varepsilon}^{(s)}(f)$ is in turn upper bounded by a constant times $D_s(f) := \|f\|_{H^s(P)}^2$.
-Together, these steps establish Proposition~\ref{prop:graph_seminorm}.
-
-When $s = 1$, similar results have been shown in \textcolor{red}{(?)}.
-On the other hand, much less work has been done in the $s > 1$ case, where the analysis is more complicated.
-For one thing, we note that when $s > 1$ the graph energy $\dotp{L^s f}{f}_n$ is itself a biased estimate of the non-local energy $E_{\varepsilon}^{(s)}(f)$.
-This bias term is the second term on the right hand side of~\eqref{eqn:graph_seminorm_1}, and is neglible only when $r \gtrsim n^{-1/(2r + d)}$.
-On the other hand, in order to give an estimate of $E_{\varepsilon}^{(s)}(f)$ in terms of the Sobolev norm $D_s(f)$, we first show that $L_{P,\varepsilon}^jf(x) \approx \Delta_P^jf(x)$, for $j = s/2$ when $s$ is odd and $j = (s - 1)/2$ when $s$ is even.
-In words, we show that a $j$th-order iterated difference operator approximates a $2j$-th (or $2j + 1$-st) order differential operator.
-In to establish this fact, we require that on the one hand that $p \in C^{r}(\mc{X};M)$, and on the other hand that $f$ be zero-trace.
-
-\textbf{(3) Lower bound on the eigenvalue.} On the other hand, recent work \textcolor{red}{(Burago14, Shi16, Garcia Trillos 18)} has analyzed the convergence of $\lambda_{k}$ towards $\lambda_{k}(\Delta_P)$.
-They provide explicit bounds on the relative error $|\lambda_{k} - \lambda_{k}(\Delta_P)|/\lambda_{k}(\Delta_P)$, showing that the relative error is small for sufficiently large $n$, small $\varepsilon$, and any $1 \leq k \leq n$ that is not too large relative to $n$.
-These results are actually stronger than are necessary to establish Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_estimation_ho}---in order to get rate-optimality, we need only show that $\lambda_{K}/\lambda_K(P)$ is $\Omega_P(1)$---but unfortunately they all assume $P$ is supported on a manifold without boundary.
-We will instead use the results of \textcolor{red}{(Green 2021)}, whose assumptions match our own, and who give a weaker bound on $\lambda_k/\lambda_k(\Delta_P)$ that will nevertheless suffice for our purposes.
-
-\begin{proposition}
+In the second step, we show that whenever $\varepsilon$  is smaller than some contanst $c$, then $E_{\varepsilon}^{(s)}(f)$ is in turn upper bounded by a constant times $M^2 \|p\|_{C^{s - 1}(P)}^2$.
+Note that neither quantity here is random; this is a purely approximation-theoretic statement.
+Together, these steps establish Proposition~\ref{prop:graph_seminorm_ho}.
+
+The analysis of $\dotp{L_{n,\varepsilon}^sf_0}{f_0}$ when $s > 1$ is substantially trickier than the first-order case, for several reasons.
+First of all, we note that when $s > 1$ the graph energy $\dotp{L_{n,\varepsilon}^s f}{f}_n$ is itself a biased estimate of the non-local energy $E_{\varepsilon}^{(s)}(f)$, in contrast to the first order case where $\Ebb\bigl[\dotp{L_{n,\varepsilon} f}{f}_n\bigr] = E_{\varepsilon}(f)$.
+In the language of classical statistics, $\dotp{L_{n,\varepsilon}^s f}{f}_n$ is a $V$-statistic, that is the sum of an unbiased estimator of $E_{\varepsilon}^{(s)}(f)$ (in other words, a $U$-statistic) plus some higher-order, pure bias terms.
+These pure bias terms are negligible only when $\varepsilon \gg n^{-1/(2r + d)}$.
+
+On the other hand, in order to give an estimate of $E_{\varepsilon}^{(s)}(f)$ in terms of the integrated, squared derivatives of $f_0$ and $p$, we first show that $L_{P,\varepsilon}^jf(x) \to \Delta_P^jf(x)$ as $\varepsilon \to 0$; here $j = (s - 1)/2$ when $s$ is odd and $j = (s - 2)/2$ when $s$ is even.
+In words, we show that a $j$th-order iterated difference operator approximates a $2j$-th order differential operator.
+In order to establish this fact, we require that on the one hand that $p \in C^{s - 1}(\mc{X})$, and on the other hand that $f$ be zero-trace.
+Crucially we do not try to show $L_{P,\varepsilon}^sf(x) \to \Delta_{P}^sf(x)$: though this may seem like a natural first step, note carefully that the latter object is an order-$2s$ differential operator, whereas we assume that $f$ has only $s$ bounded derivatives.
+
+% AG: I think this following might just be garbage, but I will leave it in in case it piques Ryan or Siva's interest.
+
+% Proposition~\ref{prop:graph_seminorm_ho}, and its proof, show that the iterated graph Laplacian $L_{n,\varepsilon}^s$ does a reasonable job of approximating a higher-order differential operator.
+The conditions we require --- that $\varepsilon$ be sufficiently large, that $f$ be zero-trace, and that $p$ have regularity of order $s - 1$ --- indicate that the iterated graph Laplacian is an imperfect, if adequate, tool for this job.
+This is no fault of the graph Laplacian.
+Rather, as pointed out by \textcolor{red}{(Sadhanala et al.
+2017)} its reflects that graph Laplacians were designed for a very general circumstance in which the only notions of derivation are first-differentials, divergences, and compositions of the two.
+Indeed, viewed in this light it is a pleasant surprise that iterated graph Laplacians approximate higher-order differentials at all.
+When one assumes Euclidean structure, as we do in this paper, there exist many methods (e.g.
+local linear embeddings, Hessian local linear embedding, local tangent space alignment) which leverage this structure to approximate differential operators in a more sophisticated manner.
+However, thus far little theoretical investigation has been done into even the pointwise behavior of these approaches.%
+
+\paragraph{Neighborhood graph eigenvalue.}
+On the other hand, several recent works \citep{burago2014,garciatrillos18,calder2019} have analyzed the convergence of $\lambda_{k}$ towards $\lambda_{k}(\Delta_P)$.
+They provide explicit bounds on the relative error $|\lambda_{k} - \lambda_{k}(\Delta_P)|/\lambda_{k}(\Delta_P)$, which show that the relative error is small for sufficiently large $n$ and small $\varepsilon$.
+Crucially, the guarantees hold simultaneously for all $1 \leq k \leq k_{\max}$, where $k_{\max}$ must satisfy $\lambda_{k_{\max}}(\Delta_P) \ll \varepsilon^{-2}$.
+These results are actually stronger than are necessary to establish Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho}---in order to get rate-optimality, we need only show that $\lambda_{K}/\lambda_K(P)$ is $\Omega_P(1)$---but unfortunately they all assume $P$ is supported on a manifold without boundary (i.e.
+they assume Model~\ref{def:model_manifold} rather than Model~\ref{def:model_flat_euclidean}).
+
+In the case where $\mc{X}$ is assumed to have a boundary, the graph Laplacian $L$ is a reasonable approximation of the operator $\Delta_P$ only at interior points $x$ such that $B(x,\varepsilon) \subseteq \mc{X}$.
+In contrast, at points $x$ near the boundary $L_{n,\varepsilon}$ is known to approximate a completely different operator altogether \citep{belkin2012}.
+This is reminiscent of the boundary effects present in the analysis of kernel smoothing.
+Thus proving convergence of $\lambda_k$ to a continuum limit becomes a substantially more challenging problem.
+We will instead use the results of \textcolor{red}{(Green et al.
+2021)}, whose assumptions match our own, and who give a weaker bound on $\lambda_k/\lambda_k(\Delta_P)$ that will nevertheless suffice for our purposes.
+
+\begin{proposition}[Lemma~2 of \textcolor{blue}{(Green et al. 2021)}]
 	\label{prop:graph_eigenvalue}
+	Suppose Model~\ref{def:model_flat_euclidean}. Then there exist constants $c$ and $C$ such that the following statement holds: if $C(\log n/n)^{1/d} < \varepsilon < c$, then
+	\begin{equation}
+	\label{eqn:graph_eigenvalue}
+	\lambda_k \geq c \cdot \min\Bigl\{\lambda_k(\Delta_P), \frac{1}{\varepsilon^{2}} \Bigr\} \quad \textrm{for all $1 \leq k \leq n$,}
+	\end{equation}
+	with probability at least $1 - Cn\exp\{-n\varepsilon^d\}$. 
+\end{proposition}
+Note immediately that $\lambda_0(\Delta_P) = \lambda_0 = 0$.
+Furthermore, Weyl's Law \textcolor{red}{(Dunlop, Wu)} tells us that under Model~\ref{def:model_flat_euclidean}, $k^{2/d} \lesssim \lambda_{k}(\Delta_P) \lesssim k^{2/d}$ for all $k \in \mathbb{N}, k > 1$.
+Combining these statements with~\eqref{eqn:graph_eigenvalue}, we conclude that $\lambda_{K} = \Omega_P(K^{2/d})$ so long as $K \lesssim \varepsilon^{-d}$.
+
+\paragraph{Empirical norm.}
 \textcolor{red}{(TODO)}
-	With probability at least $1 - \textcolor{red}{(?)}$,
-	\begin{equation*}
-	\lambda_k \geq c \cdot \min\Bigl\{\lambda_k(\Delta_P), \frac{1}{\varepsilon^{2}} \Bigr\}
-	\end{equation*}
-\end{proposition}
-By Weyl's Law \textcolor{red}{(Dunlop, Wu)}, there exist constants $c_1$ and $C_1$ which do not depend on $k$, such that $c_1 k^{2/d} \leq \lambda_{k}(\Delta_P) \leq C_1 k^{2/d}$ for all $k \in \mathbb{N}$.
-As a result, $\lambda_{K} \geq c \lambda_{K}(\Delta_P)$ as long as $\varepsilon \geq K^{-1/d}$.
-
-\textbf{(1.2) Traps we avoid.} It is tempting to use recent results regarding spectral convergence to analyze the estimator $\wh{f}$ and test $\wh{T}$.
-This suffers from an accumulation of error problem.
-
-It is also key that we directly analyze the semi-norm $\dotp{L_{n,\varepsilon}^s f}{f}_n$, rather than invoking the pointwise convergence of $L_{n,\varepsilon}^{s}f \to \Delta_P^{s}f$ in order to obtain a bound on $\dotp{L_{n,\varepsilon}^s f}{f}_n$.
-Recall that we assume $f$ has bounded derivatives up to (and including) order $s$.
-This suffices to gain control of the semi-norm $\dotp{L^s f}{f}_n$, because $\dotp{L^s f}{f}_n$ approximates the $\Leb^2$ norm of an order $s$ differential operator, and so we have exactly as many derivatives as we need.
-On the other hand, $L^s$ approximates an order $2s$ differential operator, and $\Delta_P^sf$ is only a well-defined limiting object if we assume $f$ has order $2s$ derivatives.
-
-\textcolor{red}{(TODO)}
-\begin{itemize}
-	\item Possibly turn the first part of the discussion after Proposition~\ref{prop:graph_seminorm} into a small proof.
-\end{itemize}
+
+Invoking the bounds of Propositions~\ref{prop:graph_seminorm_fo}-\ref{prop:graph_eigenvalue} inside the bias-variance tradeoffs~\eqref{eqn:estimation_biasvariance} and~\eqref{eqn:testing_biasvariance} and then choosing $K$ to balance bias and variance (if possible), leads to the conclusions of Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}.
 
@@
 
-The optimal rates for nonparametric regression suffer from a bona fide curse of dimensionality, and are thus difficult to reconcile with the practical success of many methods for nonparametric regression.
-One explanation that resolves this seeming paradox is the \emph{manifold hypothesis}, where the design points $X$ are assumed to lie on or near a manifold $\mc{X}$ of intrinsic dimension $m < d$.
-Under the manifold hypothesis nonparametric regression is demonstrably easier, and in particular it is known \citep{bickel2007,ariascastro2018} that in this setting the optimal rates over $H^s(\Xset)$ scale like $n^{-2s/(2s + m)}$ (for estimation) and $n^{-4s/(4s + m)}$ (for testing).
-
-On the other hand, a theory has been developed~\citep{belkin03,belkin05,niyogi2013} establishing the the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
-In this section, we contribute to this line of work by showing that under the manifold hypothesis, Laplacian eigenmaps achieve the sharper minimax estimation and testing rates.
-First, we give some relevant background on Riemmanian manifolds, and Sobolev classes defined upon them.
-
-\subsection{Riemmanian manifolds}
-Throughout this section, we will assume that the support $\mc{X}$ of the distribution $P$ is a manifold of dimension $m$ embedded in $\Rd$.
-We give to $\mc{X}$ the Riemannian structure induced by the ambient space $\Rd$, and denote the resulting volume form by $d\mu$.
-We will also assume some regularity conditions on $\mc{X}$ and $P$.
-\begin{enumerate}[label=(A\arabic*)]
-	\setcounter{enumi}{4}
-	\item 
-	\label{asmp:domain_manifold} The manifold $\mc{X}$ is closed, connected, smooth and boundaryless. Additionally, the injectivity radius of $\mc{X}$ is at least $i_0 > 0$.
-	\item 
-	\label{asmp:density_manifold} The distribution $P$ admits a density $p$ with respect to the volume form $d\mu$. The density is bounded away from $0$ and $\infty$,
-	\begin{equation*}
-	0 < \frac{1}{\rho} < p(x) \leq \rho < \infty
-	\end{equation*}
-	for all $x \in \mc{X}$.
-\end{enumerate}
-The definition of the Sobolev space $H^s(P)$, and the normed ball $H^s(P;M)$ are the same, mutatis mutandi, as when $\mc{X}$ is full dimensional.
+In this section we consider the manifold setting, where $(X_1,Y_1),\ldots,(X_n,Y_n)$ are observed according to Model~\ref{def:model_manifold}.
+\citet{belkin03,belkin05,niyogi2013} establish that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+We build on this work by showing that when $f_0 \in H^s(P)$ and $P$ is supported on a manifold, Laplacian eigenmaps achieve the sharper minimax estimation and testing rates reviewed in Section~\ref{subsec:minimax_rates_sobolev}.
 
 \subsection{Error rates under the manifold hypothesis}
-We now show that Laplacian eigenmaps achieves the faster minimax rates under the manifold hypothesis, when the regression function $f_0 \in H^s(P;M)$ and $s \leq 4$.
+We now show that Laplacian eigenmaps achieves the faster minimax rates under the manifold hypothesis, when the regression function $f_0 \in H^s(P;M)$, $s \leq 4$, and, in the testing problem, $m \leq 4$.
 This section will proceed in a similar fashion to Section~\ref{sec:higher_order_sobolev_classes}, except with the ambient dimension $d$ replaced by intrinsic dimension $m$.
@@
 \paragraph{Estimation.}
-To obtain a minimax optimal estimator, we choose the graph radius $\varepsilon$ and number of eigenvectors $K$ as in~\ref{asmp:parameters_estimation_ho}, except with ambient dimension $d$ replaced by the intrinsic dimension $m$.
+To ensure that $\wh{f}$ is an in-sample minimax rate-optimal estimator, we choose the graph radius $\varepsilon$ and number of eigenvectors $K$ as in~\ref{asmp:parameters_estimation_ho}, except with ambient dimension $d$ replaced by the intrinsic dimension $m$.
 
@@
 	\label{eqn:radius_estimation_manifold}
-	C_0 \cdot \max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/m}, n^{-1/(2r + m)}\biggr\} \leq \varepsilon \leq \min\{i_0, K^{-1/m}\}
+	\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/m}, n^{-1/(2(s - 1) + m)}\biggr\} \ll \varepsilon \ll \min\{1, K^{-1/m}\}
 	\end{equation}
@@
 	\begin{equation*}
-	K \asymp (M^2n)^{2m/(2s + m)}
+	K \asymp (M^2n)^{2m/(2s + m)}.
 	\end{equation*}
@@
 	\label{thm:laplacian_eigenmaps_estimation_manifold}
-	Suppose $f_0 \in H^s(\mc{X},M)$, for $s \leq 4$. Under appropriate conditions, with probability at least $1 - \delta - \textcolor{red}{(?)}$,
+	Suppose Model~\ref{def:model_manifold}, and additionally $f_0 \in H^s(P,M)$ and $p \in C^{s - 1}(\mc{X})$ for $s < 4$.
+There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_manifold}, then
 	\begin{equation}
 	\label{eqn:laplacian_eigenmaps_estimation_manifold}
-	\|\wh{f} - f_0\|_n^2 \lesssim M^{2m/(2s + m)}n^{-2s/(2s + m)}
-	\end{equation}
+	\|\wh{f} - f_0\|_n^2 \leq C\Bigl(\frac{1}{\delta}M^2(M^2n)^{-2s/(2s + m)} \wedge 1\Bigr) \vee \frac{1}{n},
+	\end{equation}
+	with probability at least $1 - \delta - Cn\exp(-cn\varepsilon^m)$.
 \end{theorem}
@@
 \paragraph{Testing.}
-To construct a minimax optimal test using $\wh{T}$, we choose $\varepsilon$ and $K$ as in~\ref{asmp:parameters_testing_fo}, except with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
+Likewise, to construct a minimax optimal test using $\wh{T}$, we choose $\varepsilon$ and $K$ as in~\ref{asmp:parameters_testing_fo}, except with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
 \begin{enumerate}[label=(A\arabic*)]
@@
 	\label{asmp:parameters_testing_manifold}
-	The graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy~\eqref{eqn:radius_estimation_manifold}, for $r = 0$. Additionally, there exist constants $c_2$ and $C_2$ such that
-	\begin{equation*}
-	c_2 (M^2n)^{2m/(4s + m)} \leq K \leq C_2 (M^2n)^{2m/(4s + m)}
+	The graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy~\eqref{eqn:radius_estimation_manifold}. Additionally, the
+	\begin{equation*}
+	K \asymp C_2 (M^2n)^{2m/(4s + m)}
 	\end{equation*}
@@
 	\label{thm:laplacian_eigenmaps_testing_manifold}
-	Suppose $f_0 \in H^1(\mc{X};M)$.
-If the Laplacian eigenmaps test $\varphi$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_testing_manifold}, then the following statement holds for any $\delta \in (0,1)$: if $f_0$ satisfies
+	Fix $a,b \in (0,1)$.
+Suppose Model~\ref{def:model_manifold}.
+Then $\mathbb{E}_0[\varphi] \leq a$, i.e $\varphi$ is a level-$a$ test.
+Suppose additionally $f_0 \in H^s(P,M)$, that $p \in C^{s-1}(\mc{X})$, and that $s < 4$ and $m \leq 4$.
+Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_testing_manifold}, and if $f_0$ satisfies
 	\begin{equation}
 	\label{eqn:laplacian_eigenmaps_testing_criticalradius_manifold}
-	\|f_0\|_P^2 \geq \frac{C}{\delta} \max\biggl\{\frac{1}{\delta}M^{2m/(4 + m)}n^{-4/(4 + m)}, M n^{-p(m)}\biggr\},
-	\end{equation}
-	then $\varphi$ has Type II error upper bounded by 
-	\begin{equation}
-	\Pbb_{f_0}(\varphi = 0) \leq \delta + \textcolor{red}{(...)}
-	\end{equation}
+	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-4s/(4s + m) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{p(m)}} \biggr) \vee \frac{1}{n},
+	\end{equation}
+	then $\Ebb_{f_0}[1 - \phi] \leq b$.
 \end{theorem}
@@
 \begin{itemize}
-	\item \textbf{Proof strategy.} The proofs of Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold} follow very similarly to the full-dimensional setting.
-The difference is that when $\mc{X}$ is a manifold with intrinsic dimension $m$, we can prove analogous results to Propositions~\ref{prop:graph_seminorm} and~\ref{prop:graph_eigenvalue}, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
-	\item \textbf{Restriction on number of derivatives.} Unlike in the full-dimensional case, our upper bound on the estimation error of $\wh{f}$ matches the minimax rate $n^{-2s/(2s + m)}$ only when $s \leq 4$.
+	\item The proofs of Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold} follow very similarly to the full-dimensional setting.
+The difference is that when $\mc{X}$ is a manifold with intrinsic dimension $m$, we can prove analogous results to Propositions~\ref{prop:graph_seminorm_fo}-\ref{prop:graph_eigenvalue}, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
+	\item Unlike in the full-dimensional case, our upper bounds on the estimation and testing error of Laplacian eigenmaps match the minimax rate only when $s < 4$.
+Our upper bounds when $s \geq 4$ follow from the embedding $H^s(P;M) \subset H^{3}(P;M)$, i.e they match the rates we get just by assuming that the $3$rd order derivative is bounded, and are clearly suboptimal.
+
 	We now explain this discrepancy.
 At a high level, thinking of the graph $G$ as an estimate of the manifold $\mc{X}$, we incur some error by using Euclidean distance rather than geodesic distance to form the edges of $G$.
-In contrast, in the full-dimensional setting the Euclidean metric exactly coincides with the geodesic distance for all points $x,z \in \mc{X}$ that are sufficiently close to each other and far from the boundary of $\mc{X}$.
-This extra error incurred in the manifold setting by using the ``wrong distance'' dominates when $s > 4$.
+This is in contrast with the full-dimensional setting, where the Euclidean metric exactly coincides with the geodesic distance for all points $x,z \in \mc{X}$ that are sufficiently close to each other and far from the boundary of $\mc{X}$.
+This extra error incurred in the manifold setting by using the ``wrong distance'' dominates when $s \geq 4$.
+
 	As this explanation suggests, by building $G$ using the geodesic distance one could avoid this error, and would obtain superior rates of convergence.
 However this is not an option for us, as we assume $\mc{X}$---and in particular its geodesics---are unknown.
-	\item \textbf{Comparison with traditional spectral series estimators.} Traditional spectral series estimators achieve the minimax rate for all values of $s$ and $m$.
-On the other hand, as we have already discussed, these estimators assume a priori knowledge of the domain $\mc{X}$, which is a particularly unrealistic assumption when $\mc{X}$ is assumed to be a manifold.
+Likewise, traditional spectral series estimators achieve the minimax rate for all values of $s$ and $m$, but are undesirable for the same reason---we do not want to assume that $\mc{X}$ is known.
 It is not clear whether this gap between spectral series and Laplacian eigenmaps estimators---or more generally, between estimators which assume the manifold is known, and those which do not---is real, or a product of loose upper bounds.
+
+	\item Finally, when $m > 4$, we get an upper bound on testing error equivalent to that of Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}, except with the ambient dimension $d$ replaced by intrinsic dimension $m$.
 \end{itemize}
@@
 \begin{itemize}
-	\item \textbf{Tuning parameters by sample-splitting.} Extending the estimator $\wh{f}$ to be defined out-of-sample gives us the option of selecting the hyper-parameters $\varepsilon$ and $K$ in a data-dependent manner---e.g.
-by sample splitting, minimizing the error over a held-out set.
-This is by far the most common practical recommendation for selecting hyper-parameters \textcolor{red}{(references)}, and addresses a serious limitation of Laplacian eigenmaps.
+	\item \textbf{Tuning parameters by sample-splitting.} Extending the estimator $\wh{f}$ to be defined out-of-sample allows us to select the hyper-parameters $\varepsilon$ and $K$ using sample splitting or cross validation.
+For instance, we can (i) split the sample into two halves, (ii) use the first half to compute $T_{h,n}\wh{f}$ for various values of $\varepsilon$, $h$, and $K$, (iii) choose the optimal values of these three hyperparameters by minimizing error on the held out set.
+This is by far the most common used way to choose hyperparameters \textcolor{red}{(references)}, and it addresses a serious limitation of Laplacian eigenmaps.
 Moreover, it is known \textcolor{red}{(references)} that choosing hyper-parameters by minimizing error over a held-out set results in estimators that optimally adapt to the order of regularity $s$---that is, that are rate-optimal (up to $\log$ factors), even when $s$ is unknown.
 We believe that similar arguments will imply that $T_{h,n}\wh{f}$ is adaptive in this sense, though we do not pursue the details.
-	\item \textbf{Other extensions.} We consider extension by kernel smoothing because it is a simple and statistically optimal procedure that does not require any knowledge of the domain $\mc{X}$ or distribution $P$.
-There exist (many) other approaches to extending (or interpolating) a function $f$ based on the evaluations $\{f(X_1),\ldots,f(X_n)\}$, and we now comment on a few of these.
+	\item \textbf{Other extensions.} We consider extension by kernel smoothing because it is a simple and statistically optimal procedure that does not require any knowledge of the domain $\mc{X}$ or distribution $P$---as we have argued, this latter property is one of the main selling points of Laplacian eigenmaps as a tool for nonparametric regression.
+
+	That being said, there exist (many) other approaches to extending (or interpolating) a function $f$ based on the evaluations $\{f(X_1),\ldots,f(X_n)\}$, and we now comment on a few of these alternatives.
 One approach is to perform minimum norm interpolation, where one defines a normed space $(\mc{F},\|\cdot\|_{\mc{F}})$ and then solves the optimization problem

### 2021-03-22 talks/laplacian_smoothing/AISTATS/short_version.tex
Added a short version of AISTATS talk.

--- 2021-03-19 talks/laplacian_smoothing/AISTATS/minimax_optimal_laplacian_smoothing.tex
@@
 
-\author{Alden Green}
+\author{Alden Green, Sivaraman Balakrishnan, Ryan Tibshirani}
 \institute{Department of Statistics and Data Science \\
@@
 \titlepage
-\centering \emph{Co-authors: Sivaraman Balakrishnan, Ryan Tibshirani}
 \end{frame}
@@
 \begin{frame}{Basic problem setup}
-\alert{Nonparametric regression, random design}: We observe $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are sampled independently from density $p$, and
+\alert{Nonparametric regression, random design}: We observe $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n \in \Reals^d$ are sampled independently from density $p$, and
 \begin{equation*}
@@
 \end{equation*}
-Goal is to learn the unknown regression function $f_0$, which is assumed to be \alert{Sobolev} smooth, i.e.
+Goal is to estimate the unknown regression function $f_0$, which is assumed to be \alert{Sobolev} smooth, i.e.
 \begin{equation*}
@@
 \end{equation*}
-Either \alert{estimate} $f_0$ with some $\wh{f}$, or \alert{test} whether $f_0 = 0$.
 \end{frame}
@@
 Laplacian smoothing solves a discrete, penalized least squares problem on a graph.
+\pause
 \begin{enumerate}
-	\item  Form a \alert{neighborhood graph} $G$ over $X_1,\ldots,X_n$, with weighted edges $W_{ij} = K(\|X_i - X_j\|/r)$. The \alert{graph Laplacian} $L$ is defined by
+	\item  Form a \alert{neighborhood graph} $G$ over $X_1,\ldots,X_n$, with weighted edges $W_{ij} = K(\|X_i - X_j\|/r)$. The \alert{graph Laplacian} $L = D - W$ induces a semi-norm,
 	\begin{equation*}
-	(Lf)_i = \sum_{i = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr) W_{ij}.
+	f^{\top} L f = \sum_{i,j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 W_{ij}.
 	\end{equation*}
+	\pause
 	\item The \alert{Laplacian smoothing} estimator is
 	\begin{equation*}
-	\wh{f} = \argmin_{f \in \Reals^n} \|Y - f\|_n^2 + \frac{\rho}{2} f^{\top} L f,\quad\textrm{with test statistic}\quad\wh{T} = \|\wh{f}\|_n^2. 
+	\wh{f} = \argmin_{f \in \Reals^n} \|Y - f\|_n^2 + \frac{\rho}{2} f^{\top} L f.
 	\end{equation*}
@@
 
-\begin{frame}{Motivation}
-Why Laplacian smoothing?
-\begin{itemize}
-	\item \emph{Computational ease.} Sparse graph, fast solvers.
-	\item \emph{Generality.} Makes sense whenever one can form a graph $G$. 
-	\item \emph{Weak supervision.} Naturally extends to semi- and unsupervised problems.
-\end{itemize}
-Intuition: when $n$ large, $r$ small and $f$ smooth, {(Bousquet et al. 2004)} show
+\begin{frame}{Why Laplacian smoothing?}
+Intuition: when $p$ uniform, $n$ large, $r$ small and $f$ smooth, {(Bousquet et al. 2004)} show
 \begin{equation*}
-f^T L f = \frac{1}{2}\sum_{i,j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 W_{i,j} \approx n^2 r^{d + 2} \int \|\nabla f\|_2^2 p^2(x) \,dx.
+f^{\top} L f = \frac{1}{2}\sum_{i,j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 W_{i,j} \approx n^2 r^{d + 2} \int \|\nabla f(x)\|_2^2 \,dx.
 \end{equation*}
-So $\wh{f}$ seems like a noisy and discrete approximation to a very classical estimator $\wt{f}$: aka \alert{smoothing spline} ($d = 1$), or \alert{thin-plate spline} ($d \geq 2$), defined as
+\pause
+So $\wh{f}$ seems like a noisy and discrete approximation to $\wt{f}$,
 \begin{equation*}
-\widetilde{f} = \argmin_{f \in H^1} \|Y - f\|_n^2 + \rho \int \|\nabla f(x)\|_2^2  \,dx.
+\widetilde{f} := \argmin_{f \in H^1} \|Y - f\|_n^2 + \rho \int \|\nabla f(x)\|_2^2  \,dx.
 \end{equation*}
-Similarly $\wh{T}$ approximates $\wt{T} = \|\wt{f}\|_n^2$. 
+Very classical estimator! Aka \alert{smoothing spline} ($d = 1$), or \alert{thin-plate spline} ($d \geq 2$).
 \end{frame}
@@
 \begin{frame}{Main Question}
-In $1d$, smoothing splines achieve the minimax optimal rate \blue{$n^{-2/3}$} (estimation) or \blue{$n^{-4/5}$} (testing) over Sobolev functions.
+In $1d$, smoothing splines achieve the minimax optimal rate \blue{$n^{-2/3}$} (estimation) over Sobolev functions.
+\pause
 \begin{block}{{\bf Question}}
@@
 \end{block}
-
+\pause
 \begin{block}{{\bf Answer}}
@@
 	\begin{itemize}
-		\item Laplacian smoothing has error of \blue{$n^{-2/3}$} (estimation) and \blue{$n^{-4/5}$} (testing) when $d = 1$, matching smoothing splines.
-		\item Laplacian smoothing also achieves optimal error \blue{$n^{-2/(2 + d)}$} (estimation) and \blue{$n^{-4/(4 + d)}$} (testing) when $d = 2,3,4$ (up to $\log n$ factor when $d = 4$).
+		\item Laplacian smoothing has error of \blue{$n^{-2/3}$}  when $d = 1$, matching smoothing splines.
+		\item Laplacian smoothing also achieves optimal error \blue{$n^{-2/(2 + d)}$} when $d = 2,3,4$ (up to $\log n$ factor when $d = 4$).
 		\item On the other hand $\wt{f}$ is \alert{not even well-defined} when $d > 1$. Well known in spline/semi-supervised learning communities.
@@
 			$d = 1$ & \blue{\bm{${n^{-2/3}}$}} & \blue{${n^{-2/3}}$} \\
-			$d = 2,3$ & \blue{\bm{${n^{-2/(2 + d)}}$}} & \red{NA}/\red{$1$} \\
-			$d = 4$ & $\bm{\blue{{n^{-1/3}}} (\log n)^{1/3}}$ & \red{NA}/\red{$1$} \\
-			$d \geq 5$  & $\bm{(\log n/n)^{4/(3d)}}$ & \red{NA}/\red{$1$} \\
-		\end{tabular}
-	\end{center}
-\end{table}
-
-{\large Testing rates, assuming $f_0 \in H^1(M)$ ($d < 4$) or $f_0 \in L^4(M)$ ($d \geq 4$)}
-\begin{table}
-	\begin{center}
-		\begin{tabular}{p{.25\textwidth} | p{.25\textwidth} p{.25\textwidth} }
-			Dimension & Laplacian smoothing & Smoothing/Thin-plate splines \\
-			\hline
-			$d = 1$ & \blue{\bm{${n^{-4/5}}$}} & \blue{${n^{-4/5}}$} \\
-			$d = 2,3$ & \blue{\bm{${n^{-4/(4 + d)}}$}} & \red{NA}/\red{$n^{-1/2}$} \\
-			$d \geq 4$ & $\blue{\bm{{n^{-1/2}}}}$ & \red{NA}/\blue{$n^{-1/2}$} \\
+			$d = 2,3$ & \blue{\bm{${n^{-2/(2 + d)}}$}} & \red{NA} \\
+			$d = 4$ & $\bm{\blue{{n^{-1/3}}} (\log n)^{1/3}}$ & \red{NA} \\
+			$d \geq 5$  & $\bm{(\log n/n)^{4/(3d)}}$ & \red{NA} \\
 		\end{tabular}
@@
 
-\begin{frame}{Explanation: Failure of Thin-plate Splines}
-Well-posedness of smoothing splines (when $d = 1$) vs. ill-posedness of thin-plate splines ($d \geq 2$) explained by \alert{Sobolev Embedding Theorem}.
+\begin{frame}{Extensions}
+Several different directions.
 \begin{itemize}
-	\item In $1d$, every function in $H^1$ is continuous. 
-	\item When $d \geq 2$, this is no longer the case.
+	\item \alert{Goodness-of-fit testing}: Use statistic $\wh{T} = \|\wh{f}\|_n^2$ to test whether $f_0 = 0$. Do we achieve optimal rates \blue{$n^{-4/(4 + d)}$}? \blue{(Yes! See paper.)}
+	\item \alert{Manifold adaptivity}: When $p$ has support on a manifold of dimension $m < d$, optimal rates are \blue{$n^{-2/(2 + m)}$} (estimation) and \blue{$n^{-4/(4 + m)}$} (testing).
+Does Laplacian smoothing automatically adapt?
+\blue{(Yes!
+See paper.)}
+	\item \alert{Out-of-sample estimation}: Can we extend estimator $\wh{f}$ to have optimal out-of-sample mean squared error? \blue{(Yes, if $f_0$ is Lipschitz! See paper.)}
 \end{itemize}
-As a result the criterion underlying $\wt{f}$ is \alert{not continuous} with respect to $H^1$ norm. (Green and Silverman 1994) construct sequences of ``bump functions'' $f_1,f_2,\ldots$ such that
-\begin{equation*}
-f_k(X_i) = Y_i \quad \text{for all $i = 1,\ldots,n$, but} \quad \|f_k\|_{H^1} \to 0.
-\end{equation*}
-\alert{Bottom line}: The continuum thin-plate spline problem assigns (much) higher loss + penalty to the regression function $f_0$ than to bump functions $f_k$. \red{\frownie} 
 \end{frame}
 
-\begin{frame}{Explanation: Success of Laplacian Smoothing}
-\alert{Discretization} is a blessing!  
-\begin{itemize}
-	\item Previous construction of ``bump functions'' $f_1,f_2,\ldots$ will satisfy
+\begin{frame}[t]{Summary}
+	Specific takeaway: Laplacian smoothing is an optimal method for nonparametric regression. \newline
+
+	General takeaway: Discretization can allow us to solve problems even when continuum analogues are not well-posed.
+	\newline 
+	\newline
+	\newline
+	\pause
+	\begin{center}
+	{\huge Thank You!}
+	\end{center}
+
+\end{frame}
+
+\begin{frame}{Analysis}
+We use a conditional-on-design \alert{bias-variance} decomposition, based on the eigenvalues $\lambda_k$ of $L$. \newline
+
+For estimation,
 \begin{equation*}
-	f_k^{\top} L f_k = \sum_{i,j = 1}^{n} (Y_i - Y_j)^2 K(\|X_i - X_j\|/r) \asymp n^2r^{d}.
+\|\wh{f} - f_0\|_n^2 \lesssim \rho \frac{f_0^{\top} L f_0}{n} + \frac{1}{n} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2}.
 \end{equation*}
-	\item If $f_0$ is Sobolev smooth, then $f_0^{\top} L f_0 \lesssim n^2r^{d + 2}$.
-	\item Thus for $\lambda \gg (nr^{d + 2})^{-1} r^{-2}$, we have
+
+For testing, level-$\alpha$ test $\varphi = 1\{\wh{T} \geq t_{\alpha}\}$ has negligible Type II error when
 \begin{equation*}
-	\|Y - f_k\|_n^2 + \rho f_k^{\top} L f_k \gg \|Y - f_0\|_n^2 + \rho f_0^{\top} L f_0.
+\|f_0\|_n^2 \gtrsim \rho \frac{f_0^{\top} L f_0}{n} + \frac{1/\sqrt{\alpha}}{n} \sum_{k = 1}^{n} \sqrt{\frac{1}{(\rho \lambda_k + 1)^4}}.
 \end{equation*}
-\end{itemize}
-\alert{Bottom line}: the discrete problem assigns (much) higher loss + penalty to bump functions $f_k$ than to regression function $f_0$. \blue{\smiley}
+
+Then we make use of some recent results (Burago et al. 2014, Garcia Trillos et al. 2019) on convergence of eigenvalues $\lambda_k$, and derive some of our own.
 \end{frame}
@@
 
-\begin{frame}{Extensions}
-Several different directions.
+\begin{frame}{Motvation (ctd.)}
+Why Laplacian smoothing?
 \begin{itemize}
-	\item \alert{Manifold adaptivity}: When $p$ has support on a manifold of dimension $m < d$, optimal rates are \blue{$n^{-2/(2 + m)}$} (estimation) and \blue{$n^{-4/(4 + m)}$} (testing).
-Does Laplacian smoothing automatically adapt?
-\blue{(Yes!
-See paper.)}
-	\item \alert{Out-of-sample estimation}: Can we extend estimator $\wh{f}$ to have optimal out-of-sample mean squared error? \blue{(Yes, if $f_0$ is Lipschitz! See paper.)}
-	\item \alert{Smoothness adaptivity}: When $f_0$ has $s$ smooth derivatives, optimal rates are \blue{$n^{-2s/(2s + m)}$} (estimation) and \blue{$n^{-4s/(4s + m)}$} (testing).
-Does Laplacian smoothing adapt?
-(Future work...)
-	\item \alert{Different methods}: Spectral cutoff? $k$NN graph? Normalized Laplacian? (Future work...)
+	\item \emph{Computational ease.} Sparse graph, fast solvers.
+	\item \emph{Generality.} Makes sense whenever one can form a graph $G$. 
+	\item \emph{Weak supervision.} Naturally extends to semi- and unsupervised problems.
 \end{itemize}
-\end{frame}
-
-\begin{frame}[t]{Summary}
-	Specific takeaway: Laplacian smoothing is an optimal method for nonparametric regression. \newline
-
-	General takeaway: Discretization can allow us to solve problems even when their continuum analogues are not well-posed.
-	\newline 
-	\newline
-	\newline
-
-	\begin{center}
-	{\huge Thank You!}
-	\end{center}
-
-\end{frame}
-
-\begin{frame}{Analysis}
-We use a conditional-on-design \alert{bias-variance} decomposition, based on the eigenvalues $\lambda_k$ of $L$. \newline
-
-For estimation,
-\begin{equation*}
-\|\wh{f} - f_0\|_n^2 \lesssim \frac{f_0^{\top} L f_0}{n\lambda_k} + \frac{1}{n} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2}.
-\end{equation*}
-
-For testing, level-$\alpha$ test $\varphi = 1\{\wh{T} \geq t_{\alpha}\}$ has negligible Type II error when
-\begin{equation*}
-\|f_0\|_n^2 \gtrsim \frac{f_0^{\top} L f_0}{n\lambda_k} + \frac{1}{n} \sum_{k = 1}^{n} \sqrt{\frac{1}{(\rho \lambda_k + 1)^4}}.
-\end{equation*}
-
-Then we make use of some recent results (Burago et al. 2014, Garcia Trillos et al. 2019) on convergence of eigenvalues $\lambda_k$, and derive some of our own.
+\pause
 \end{frame}

### 2021-03-23 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
All sections of Laplacian eigenmaps manuscript passed over.

@@
 \usepackage{multirow}
+\usepackage[font={small,it}]{caption}
 
@@
 
-Suppose we observe data $X_1,\ldots,X_n$ sampled independently from an unknown distribution $P$.
-As a replacement for $P$, we form a geometric graph $G$ over the observed data, with weighted edges $W_{ij}$ corresponding to proximity between samples $X_i$ and $X_j$.
-Geometric graphs encode information about $P$--- for instance, the number of connected components of its support, and the shape of each connected component---in an extremely general manner, and can be leveraged to conduct many different fundamental statistical tasks.
-These include clustering, manifold learning, semi-supervised learning, classification, and regression---both estimation and goodness-of-fit testing.
-Though much theoretical work has been done on the consistency of learning methods defined with respect to geometric graphs---and more recently, rates of convergence have been established for some methods---little is known so far about their optimality, even for classic statistical tasks.
-
-In this paper we focus on the theoretical statistical properties of \textcolor{red}{graph-based regression}.
+Suppose we observe data $X_1,\ldots,X_n \in \Reals^d$, sampled independently from an unknown distribution $P$.
+As a replacement for $P$, we form a geometric graph $G$ over the observed data, with vertices at $X_1,\ldots,X_n$ and weighted edges $W_{ij}$ corresponding to proximity between samples $X_i$ and $X_j$.
+Geometric graphs encode both local information and global information about $P$ in an extremely general manner.
+For this reason they have been leveraged to conduct many different fundamental statistical tasks, such clustering, manifold learning, semi-supervised learning, classification, and regression \textcolor{red}{(TODO): references}.
+Substantial theoretical progress has been made regarding the consistency of \emph{graph-based learning}, that is, learning algorithms defined with respect to geometric graphs \textcolor{red}{(TODO): references}.
+This work sheds light on why such procedures work, by showing that they converge to interesting \emph{continuum} limits as $n \to \infty$.
+However, thus far little has been said regarding the optimality of graph-based learning methods, even for classic statistical tasks.
+
+In this paper we focus on the theoretical statistical properties of a particularly popular graph-based learning method for regression.
 We assume that in addition to the design points $X_1,\ldots,X_n$ one observes real-valued responses $Y_1,\ldots,Y_n$, and seeks to learn the unknown regression function $f_0(x) := E[Y|X = x]$.
-The methods we will study are based on \emph{Laplacian eigenmaps}, first introduced by \cite{belkin03a}, which takes the eigenvectors of the graph Laplacian $L$.
-The Laplacian $L$ is a difference operator, acting on functions $f: \{X_1,\ldots,X_n\} \to \Reals$ as follows,
+The specific graph-based method we study is \emph{Laplacian eigenmaps}, first introduced by \cite{belkin03a}, which projects the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto the span of eigenvectors of a graph Laplacian.
+We focus on the the unnormalized graph Laplacian $L$, which is a difference operator acting on functions $f: \{X_1,\ldots,X_n\} \to \Reals$ as follows,
 \begin{equation}
@@
 \end{equation}
-The Laplacian $L$ can be seen as a discretization of the Laplace-Beltrami operator $\Delta_P$, defined as
+The Laplacian $L$ can be seen as a discretization of a weighted Laplace-Beltrami operator $\Delta_P$, defined when $P$ admits a density $p$ as
 \begin{equation}
 \label{eqn:fokker_planck_1}
-\Delta_Pf= \frac{-1}{p} \mathrm{div}(p^2 \nabla f) = \Delta f - 2\frac{\nabla p^{\top} \nabla f}{p}.
-\end{equation}
-The eigenvectors of $L$ form an orthonormal basis of $L^2(X)$, and are approximations to the eigenvectors of $\Delta_P$.
-Their corresponding eigenvalues are estimates of~\textcolor{red}{(...)}, and provide a notion of smoothness for each eigenvector---the smaller the eigenvalue, the smoother the corresponding eigenvector.
-
-Laplacian eigenmaps is thus an example of a \emph{spectral series} estimator.
-A spectral series estimator is a special case of an orthogonal series estimator, one of the most classical methods of non-parametric regression.
-Traditionally, orthogonal series estimators are defined with respect to a fixed set of basis functions, orthogonalized in $L^2(Q)$ with respect to some pre-determined reference measure $Q$.
+\Delta_Pf= -\frac{1}{p} \mathrm{div}(p^2 \nabla f).
+\end{equation}
+The eigenvectors of $L$ form an orthonormal basis of $L^2(P_n)$, and are approximations to the eigenfunctions of $\Delta_P$.
+Their corresponding eigenvalues are likewise estimates of eigenvalues of $\Delta_P$, and thus give a notion of smoothness to each eigenvector---roughly speaking, the smaller the eigenvalue, the smoother the corresponding eigenvector.
+
+Laplacian eigenmaps is thus a type \emph{spectral series} method.
+Spectral series methods, and more general orthogonal series methods, are some of most classical methods for non-parametric regression.
+Traditionally, orthogonal series methods are defined with respect to a fixed set of basis functions, orthogonalized in $L^2(Q)$ with respect to some pre-determined reference measure $Q$.
 Spectral series estimators take a particular orthogonal basis, the eigenfunctions of $\Delta_Q$.
-In contrast, eigenvectors of the graph Laplacian are data-dependent objects, and adapt to the geometry of the design distribution $P$ in a rich manner.
-
-The theoretical properties of (classical) spectral series methods are various, and are in general extremely well-understood \textcolor{red}{(references)}.
+In contrast, eigenvectors of the graph Laplacian are data-dependent objects, and adapt to the geometry of the unknown design distribution $P$ in a rich manner.
+
+The theoretical properties of (classical) spectral series methods are various, and are in general extremely well-understood \textcolor{red}{(TODO): references}.
 In particular, they are minimax rate optimal for nonparametric regression over H\"{o}lder and Sobolev spaces.
-However, in practice these methods suffer from some serious drawbacks.
-As a practical matter,  even finding the eigenfunctions of $\Delta_Q$ is in general non-trivial.
-The statistical optimality of such estimators also hinges on the reference measure $Q$ being equal to $P$, an unrealistic assumption when $P$ is unknown.
-If $Q \neq P$, the standard recommendation is to instead perform nonparametric least squares---this can be statistically rate-optimal, but is numerically unstable when the dimension $d$ is even moderately large.
-
-\textcolor{red}{(TODO)}: Change the preceding paragraph.
-
-Laplacian eigenmaps avoids these drawbacks \textcolor{red}{(...)} On the other hand its statistical properties are not as well understood \textcolor{red}{(...)}. Thus, there exists a theoretical gap.
+However, these methods suffer from some serious drawbacks.
+As a practical matter, even computing the eigenfunctions of $\Delta_Q$ is in general non-trivial.
+Moreover, the statistical optimality of such estimators also requires that the reference measure $Q$ be equal to $P$, whereas in this work we shall always assume $P$ is unknown.
+Of course there exist workarounds for these issues, for instance non-parametric least squares or \textcolor{red}{(...)}.
+We will comment more on these alternatives later; for now we merely emphasize that they constitute a fundamental change to the method under consideration.
+\textcolor{red}{Last sentence is weak, need to clarify, can maybe discuss with Ryan + Siva.}
+
+In contrast, Laplacian eigenmaps seeks to directly approximate a orthogonal series method that uses the eigenfunctions of $\Delta_P$, as we elaborate on momentarily.
+Importantly, it avoids the (arguably unrealistic) assumption that $P$ and the eigenfunctions of $\Delta_P$ be known, by using the graph Laplacian eigenvectors as a stand-in.
+On the other hand its theoretical statistical properties, and in particular its optimality as a method for regression, remain poorly understood.
 
@@
 \begin{quote}
-	\textcolor{red}{(TODO)}
+	Is Laplacian eigenmaps an optimal method for nonparametric regression over Sobolev functions?
 \end{quote}
 
-Broadly speaking, we show that when the regression function $f_0$ is smooth, in the Sobolev sense of having weak derivatives bounded in $\Leb^2$-norm, Laplacian eigenmaps methods are statistically minimax optimal, for both estimation and goodness-of-fit testing.
-This statement holds for different relations between the dimension $d$ and number of derivatives $s$, depending on the problem (estimation or testing).
+Broadly seeking, our answer is yes.
+We consider two different data models, one of which assumes a full-dimensional, flat Euclidean domain, and the other of which assumes that the distribution $P$ is supported on a manifold of small intrinsic dimension $m$.
+We show that when the regression function $f_0$ is smooth, in the Sobolev sense of having weak derivatives bounded in $\Leb^2$ norm, Laplacian eigenmaps methods are statistically minimax optimal, for both estimation and goodness-of-fit testing.
+Our statements holds for different relations between the dimension $d$ (or $m$) and number of derivatives $s$, depending on the problem (estimation or testing), and are summarized in Tables~\ref{tbl:estimation_rates} and~\ref{tbl:testing_rates}.
 \begin{table}
@@
 Bold font marks minimax optimal rates.
-In each case, rates hold for all $d \in \mathbb{N}$ (under Model~\ref{def:model_flat_euclidean}), and for all $d, m \in \mathbb{N}$ (under Model~\ref{def:model_manifold}).
-Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $M \gtrsim n^{-1/2}$.}
+In each case, rates hold for all $d \in \mathbb{N}$ (under Model~\ref{def:model_flat_euclidean}), and for all $m \in \mathbb{N}, 1 < m < d$ (under Model~\ref{def:model_manifold}).
+Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps estimator is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $M \gtrsim n^{-1/2}$.}
 	\label{tbl:estimation_rates}
@@
 	\end{center}
-	\caption{Summary of testing rates over Sobolev balls.
+	\caption{Summary of Laplacian eigenmaps testing rates over Sobolev balls.
 Bold font marks minimax optimal rates.
 Rates when $d > 4s$ assume that $f_0 \in L^4(P,M)$.
-Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $M \gtrsim n^{-1/2}$.}
+Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $M \gtrsim n^{-1/2}$.}
 	\label{tbl:testing_rates}
@@
 \paragraph{Related Work.}
+\textcolor{red}{(TODO)}: Flesh this out.
 \begin{itemize}
-	\item \textcolor{red}{(TODO)} Summarize what is known on fixed graphs.
-	\item \textcolor{red}{von Luxburg, Belkin, Slepcev} show that for a fixed index $k$, the eigenvector $v_k$ converges to the eigenfunction $\psi_k$, in the sense that \textcolor{red}{(...)}.
+	\item Summarize what is known on fixed graphs, drawing from our paper on Laplacian smoothing.
+	\item \textcolor{red}{von Luxburg, Belkin, Slepcev} show that for a fixed index $k$, the eigenvector $v_k$ converges to a eigenfunction $\psi_k$ of the Laplace-Beltrami operator $\Delta_P$.
 \textcolor{red}{Shi, Burago, Garcia Trillos, Calder, Cheng} build on this, giving finite sample bounds, rates of convergence, and making statements uniform over $1 \leq k \leq k_{\max}$.
-This suggests that Laplacian eigenmaps may be good alternatives to classical spectral series methods when the latter cannot be implemented (e.g.
-because $P$ or $\mc{X}$ are unknown.)
+This makes it clear that Laplacian eigenmaps is approximating a classical spectral series method, and may be a good alternative when the latter cannot be implemented (e.g.
+because $P$ or $\mc{X}$ are unknown.) It is tempting to try to directly use these results along with the triangle inequality to deduce that if (i)  spectral series methods are optimal, and (ii) Laplacian eigenmaps is converging to a spectral series method, then (iii) Laplacian eigenmaps must be also optimal.
+Unfortunately, this won't work, both because we will need to take an increasing number of eigenvectors (increasing, that is, with the sample size $n$) to achieve an optimal estimator and test, and because the best known rates of convergence of $v_k \to \psi_k$ are slower than the minimax rates.
 	\item There has been limited analysis of the estimator $\wh{f}$.
-\cite{zhou2011} consider a similar estimator, but in the semi-supervised setting where the number of unlabeled points grows to infinity.
-In this case the estimator reduces to a classical spectral series estimator, and there is no error incurred by approximating $\Delta_P$.
+\cite{zhou2011} consider $\wh{f}$, but in the semi-supervised setting where the number of unlabeled points grows to infinity.
+In this case $\wh{f}$ reduces to a classical spectral series estimator, and there is no error incurred by approximating $\Delta_P$.
 \cite{lee2016} consider the \emph{diffusion maps} estimator---which uses the eigenvectors of a different normalization of the Laplacian $L$---in both the supervised and semi-supervised setups, but they derive suboptimal rates for the supervised problem.
 As far as we know, there has been no analysis of the test statistic $\wh{T}$ in the random design framework which we study.
-	\item \textcolor{red}{(TODO)}: Alternatively, \cite{trillos2020} \textcolor{blue}{(Green 2021)} use the graph Laplacian operator to induce a penalty over functions $f: \{\bf X\} \to \Reals$.
+	\item Alternatively, \cite{trillos2020} \textcolor{blue}{(Green 2021)} use the graph Laplacian operator to induce a penalty over functions $f: \{\bf X\} \to \Reals$.
 The \emph{Laplacian smoothing} estimator $\wh{f}_{\mathrm{LS}}$ is obtained by minimizing the sum of this penalty along with a data-fidelity term,
@@
 
-\textbf{(5) Organization.}
-
-\textcolor{red}{(TODO)}
-
-\textbf{(6) Notation.}
+\paragraph{Organization.}
+
+We now outline the structure of the rest of this paper.
+In Section~\ref{sec:setup_main_results}, we formally define the regression problem and estimator we consider, and give some background about minimax regression over Sobolev spaces.
+In Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity}, we establish our main results regarding the optimality of Laplacian eigenmaps.
+The Laplacian eigenmaps estimator $\wh{f}$ is defined only at the design points $X_1,\ldots,X_n$.
+In Section~\ref{sec:out_of_sample}, we propose a method for out-of-sample extension of $\wh{f}$, and show that it has optimal out-of-sample error.
+In Section~\ref{sec:experiments} we examine the empirical behavior of Laplacian eigenmaps, and compare it to some natural competitors for non-parametric regression.
+We conclude with some discussion in Section~\ref{sec:discussion}.
+
+\paragraph{Notation.}
 
@@
 \begin{itemize}
-	\item Define $\Leb^2(\mc{X}), \Leb^2(P)$ and $\Leb^2(P_n)$. Define $C^k(\mc{X})$, its norm, and its normed ball.
+	\item Define Lebesgue measure $\nu$. 
+	\item Define Euclidean norm $\|\cdot\|$.
+	\item Define $\Leb^q(\mc{X}), \Leb^2(P)$ and $\Leb^2(P_n)$. Define $C^k(\mc{X})$, and its norm. Define $C_0^k(\mc{X})$.
 	\item Define asymptotic notation: $\ll$, $\lesssim$ and $\asymp$.
@@
 
-In the following, recall that the injectivity radius of a $m$-dimensional Riemannian manifold $\mc{X}$ is the maximum value of $\delta$ such that the exponential map $\exp_x: B_m(0,\delta) \subset T_x(\mc{X}) \to B_{\mc{X}}(x,\delta) \subset \mc{X}$ is a diffeomorphism for all $x \in \mc{X}$.
+In the following, we recall that the injectivity radius of a $m$-dimensional Riemannian manifold $\mc{X}$ is the maximum value of $\delta$ such that the exponential map $\exp_x: B_m(0,\delta) \subset T_x(\mc{X}) \to B_{\mc{X}}(x,\delta) \subset \mc{X}$ is a diffeomorphism for all $x \in \mc{X}$.
 \begin{definition}[Manifold model]
@@
 The injectivity radius of $\mc{X}$ is upper bounded by a constant $i_0$.
-The design distribution $P$ admits a density $p$ with respect to the volume form $d\mu$ induced by the Riemannian structure, which is bounded away from $0$ and $\infty$,
+The design distribution $P$ admits a density $p$ with respect to the volume form $d\mu$ induced by the Riemannian structure of $\mc{X}$, which is bounded away from $0$ and $\infty$,
 	\begin{equation*}
@@
 Both are derived from eigenvectors of the neighborhood graph Laplacian operator.
-For a positive, symmetric kernel $\eta: [0,\infty) \to [0,\infty)$, and a bandwidth parameter $\varepsilon > 0$, the \emph{neighborhood graph Laplacian} operator $L_{n,\varepsilon}$ is
+For a positive, symmetric kernel $\eta: [0,\infty) \to [0,\infty)$, and a bandwidth parameter $\varepsilon > 0$, the \emph{neighborhood graph Laplacian} operator $L_{n,\varepsilon}$ is defined as
 \begin{equation}
@@
 
-On a different note, it is worth explicitly pointing out that $\wh{f}$ is defined only in-sample, that is, only at the design points $X_1,\ldots,X_n$.
+% AG: Said elsewhere. 
+% On a different note, it is worth explicitly pointing out that $\wh{f}$ is defined only in-sample, that is, only at the design points $X_1,\ldots,X_n$.
 There is no straightforward or agreed upon way to extend $\wh{f}$ out-of-sample, and we address the problem of out-of-sample estimation in some detail later in Section~\ref{sec:out_of_sample}.
+%
 
@@
 Take $\mc{X}$ to be an open, connected set with Lipschitz boundary, as in Model~\ref{def:model_flat_euclidean}.
-Let $Q$ be a measure supported on a subset of $\mc{X}$, that is absolutely continuous with respect to the full-dimensional Lebesgue measure $\nu$.
-Recall that for a given multiindex $\alpha \in \mathbb{N}^d$, a function $f$ is \emph{$\alpha$ weakly differentiable} if there exists some $h \in L^1(Q)$ such that
+Recall that for a given multiindex $\alpha \in \mathbb{N}^d$, a function $f$ is \emph{$\alpha$-weakly differentiable} if there exists some $h \in L^1(\mc{X})$ such that
 \begin{equation*}
@@
 If such a function $h$ exists, it is the $\alpha$th weak partial derivative of $f$, and denoted by $D^{\alpha}f = h$.
-If $f$ is $\alpha$ weakly differentiable for all $|\alpha| = j$, then $d^jf := \sum_{|\alpha| = j}D^{\alpha}f$ is the sum of all weak partial derivatives of order $j$.
-
-For an integer $s \geq 1$, the Sobolev class $H^s(Q)$ consists of all those functions $f \in \Leb^2(Q)$ which have weak derivatives of all orders $j = 1,\ldots,s$, all of which also belong to $\Leb^2(Q)$.
-The $j$th order semi-norm for $f \in H^s(Q)$ is $|f|_{H^j(Q)} := \|d^jf\|_{L^2(Q)}$.
-The corresponding norm
-\begin{equation*}
-\|f\|_{H^s(Q)}^2 := \|f\|_{\Leb^2(Q)}^2 + \sum_{j = 1}^{s} |f|_{H^j(Q)}^2,
-\end{equation*}
-which induces the ball
-\begin{equation*}
-H^s(Q;M) := \bigl\{f \in H^s(Q): \|f\|_{H^s(Q)} \leq M\bigr\}.
-\end{equation*} 
-In the special case where $Q$ is the uniform measure over $\mc{X}$, we use the notation $H^s(\mc{X}) := H^s(Q)$.
-Finally, we note that $H^s(Q)$ is the completion of $C^{(\infty)}(\mc{X})$ with respect to the $\|\cdot\|_{H^s(Q)}$ norm, so that $C^{\infty}(\mc{X})$ is dense in $H^s(\mc{X})$.
+If $f$ is $\alpha$-weakly differentiable for all $|\alpha| = j$, then $d^jf := \sum_{|\alpha| = j}D^{\alpha}f$ is the sum of all weak partial derivatives of order $j$.
+
+For an integer $s \geq 1$, the Sobolev class $H^s(\mc{X})$ consists of all those functions $f \in \Leb^2(\mc{X})$ which have weak derivatives of all orders $j = 1,\ldots,s$, all of which also belong to $\Leb^2(\mc{X})$.
+The $j$th order semi-norm for $f \in H^s(\mc{X})$ is $|f|_{H^j(\mc{X})} := \|d^jf\|_{L^2(Q)}$, and the corresponding norm
+\begin{equation*}
+\|f\|_{H^s(\mc{X})}^2 := \|f\|_{\Leb^2(Q)}^2 + \sum_{j = 1}^{s} |f|_{H^j(Q)}^2,
+\end{equation*}
+induces a ball
+\begin{equation*}
+H^s(\mc{X};M) := \bigl\{f \in H^s(\mc{X}): \|f\|_{H^s(\mc{X})} \leq M\bigr\}.
+\end{equation*} 
+We note that $H^s(\mc{X})$ is the completion of $C^{\infty}(\mc{X})$ with respect to the $\|\cdot\|_{H^s(\mc{X})}$ norm, so that $C^{\infty}(\mc{X})$ is dense in $H^s(\mc{X})$.
 
 \paragraph{Manifold case.}
-
-\textcolor{red}{(TODO)}
+There are several ways to define Sobolev spaces on a manifold \textcolor{red}{(references)}, but we will stick with a simple definition that suffices when $\mc{X}$ is a smooth Riemmanian manifold without boundary, as in Model~\ref{def:model_manifold}.
+Let $\{\phi_{\beta}, U_{\beta}\}$ be a $C^{\infty}$ atlas of $\mc{X}$, and let $\{h_\beta\}$ be a partition of unity subordinate to $\{(\phi_{\beta}), U_{\beta}\}$.
+For a function $f \in C^{\infty}(\mc{X})$, define the Sobolev norm to be,
+\begin{equation*}
+\|f\|_{H^s(\mc{X})}^2 = \sum_{\beta} \|(fh_{\beta}) \circ \phi_\beta^{-1}\|_{H^{k}(\Reals^m)}^2.
+\end{equation*}
+Then the order-$s$ Sobolev space $H^s(\mc{X})$ is defined to be the completion of $C^{\infty}(\mc{X})$ with respect to the norm $\|\cdot\|_{H^s(\mc{X})}$.
 
@@
 In the flat Euclidean case, when $s > 1$ we will assume that $f_0$ satisfies some boundary conditions.
-In particular, we will assume that $f_0$ belongs to the set of \emph{zero trace} Sobolev functions $H_0^{s}(P)$, defined as the closure of $C_0^s(\mc{X})$ with respect to $\|\cdot\|_{H^s(P)}$ norm.
+In particular, we will assume that $f_0$ belongs to the set of \emph{zero trace} Sobolev functions $H_0^{s}(P)$, defined as the closure of $C_0^s(\mc{X})$ with respect to $\|\cdot\|_{H^s(\mc{X})}$ norm.
 Of course, under Model~\ref{def:model_manifold} the domain $\mc{X}$ is without boundary, and so boundary conditions are irrelevant.
@@
 where $\partial/(\partial {\bf n})$ is the partial derivative operator in the direction of the normal vector $\mathbf{n}$ to the boundary of $\mc{X}$.
-We define an ellipsoid $\Theta \subset \ell^2(\mathbb{N})$  using the eigenvalues $\lambda_1(\Delta_P),\lambda_2(\Delta_P),\ldots$ as $\Theta := \{\theta \in \ell^2(\mathbb{N}): \sum_{k = 1}^{\infty} \lambda_k(\Delta_P)^{2s} \theta_k^2 < \infty\}$.
-Then, the function class $\mc{H}^s(P) = \{f = \sum_{k = 1}^{\infty} \theta_k \psi_k: \theta \in \Theta\}$ consists precisely of those Sobolev functions $f \in H^s(P)$ which satisfy the following \emph{Neumann} boundary conditions,
+We define the ellipsoid $\Theta \subset \ell^2(\mathbb{N})$  using the eigenvalues $\lambda_1(\Delta_P),\lambda_2(\Delta_P),\ldots$ to be
+\begin{equation*}
+\Theta := \Bigl\{\theta \in \ell^2(\mathbb{N}): \sum_{k = 1}^{\infty} [\lambda_k(\Delta_P)]^{2s} \theta_k^2 < \infty\Bigr\}
+\end{equation*}
+Then, the function class $\mc{H}^s(P) = \{f = \sum_{k = 1}^{\infty} \theta_k \psi_k: \theta \in \Theta\}$ consists precisely of those Sobolev functions $f \in H^s(\mc{X})$ which satisfy the following \emph{Neumann} boundary conditions,
 \begin{equation}
@@
 As we have already reviewed, the eigenvectors $v_k$ of the graph Laplacian $L_{n,\varepsilon}$ converge to $\psi_k$ as $n \to \infty, \varepsilon \to 0$.
-As a result, is natural that they accurately approximate only those functions $f_0 \in \mc{H}^s(P) \subset H^s(P)$.
+Naturally, then, these eigenvectors accurately approximate only those functions $f_0 \in \mc{H}^s(P) \subset H^s(\mc{X})$.
 Although the zero-trace boundary condition is more restrictive than~\eqref{eqn:condition}---since it also requires that derivatives of even-order be equal to $0$---the point is that some kind of boundary condition will be necessary in order to obtain optimal rates for spectral series methods, including Laplacian eigenmaps.
@@
 \end{equation}
-here the infimum is taken over all estimators $\wh{f}$, the supremum over all $f_0 \in H^1(\mc{X};M)$ (first-order case) or $f_0 \in H_0^s(\mc{X};M)$ (higher-order case).
+here the infimum is taken over all estimators $\wh{f}$, and the supremum over all $f_0 \in H^1(\mc{X};M)$ (first-order case) or $f_0 \in H_0^s(\mc{X};M)$ (higher-order case).
 
 The upper bound in~\eqref{eqn:sobolev_space_minimax_estimation_rate} can be certified by a particular spectral series estimator.
-Under the assumptions of Model~\ref{def:model_flat_euclidean}, the eigenvalues $\lambda_k(\Delta_P)$ satisfy the Weyl's Law asymptotic scaling $\lambda_{k}(\Delta_P) \asymp k^{2/d}$, a fact that we discuss in more detail later.
-A key consequence of this scaling is that the Sobolev ball $H_0^s(P) \subset \mc{H}^s(P)$ continuously embeds into the Sobolev ellipsoid $\mc{\wt{H}}^s(P) = \{f = \sum_{k = 1}^{\infty} \theta_k \psi_k: \sum_{k = 1}^{\infty} (k - 1)^{2s/d} \theta_k^2\}$.
+Under the assumptions of Model~\ref{def:model_flat_euclidean}, the eigenvalues $\lambda_k(\Delta_P)$ satisfy the Weyl's Law asymptotic scaling $\lambda_{k}(\Delta_P) \asymp k^{2/d}$, an important fact that we will make use of and discuss in more detail later on.
+A key consequence of this scaling is that the Sobolev ball $H_0^s(\mc{X}) \subset \mc{H}^s(P)$ continuously embeds into the Sobolev ellipsoid $\mc{\wt{H}}^s(P) = \{f = \sum_{k = 1}^{\infty} \theta_k \psi_k: \sum_{k = 1}^{\infty} (k - 1)^{2s/d} \theta_k^2\}$.
 By reducing the regression problem to an estimation problem in a Gaussian sequence model \textcolor{red}{(see e.g.
-Tsybakov,Johnstone,Gine)}, it can be shown the estimator
+Tsybakov, Johnstone, Gine)}, it can be shown the estimator
 \begin{equation}
@@
 \end{equation}
-achieves the minimax rate over $\mc{H}^s(P;M)$, and the same upper bound on the minimax rate therefore applies to $H_0^s(P;M)$.
+achieves the minimax rate over normed balls in $\mc{\wt{H}}^s(P)$, and the same upper bound on the minimax rate therefore applies to normed balls $H_0^s(\mc{X};M)$ as well.
 
@@
 \begin{equation*}
-\epsilon(\mc{F}) := \inf\Bigl\{\epsilon > 0: \inf_{\phi} R_n(\phi,\mc{F},\epsilon) \leq b \Bigr\},
+\epsilon_n(\mc{F}) := \inf\Bigl\{\epsilon > 0: \inf_{\phi} R_n(\phi,\mc{F},\epsilon) \leq b \Bigr\},
 \end{equation*} 
@@
 \label{eqn:sobolev_space_testing_critical_radius}
-\epsilon^2\bigl(\mc{H}^s(P;M)\bigr) \asymp M^2(M^2n)^{-4s/(4s + d)}~~\textrm{for $1 \leq d < 4s$,}
+\epsilon_n^2\bigl(\mc{H}^s(P;M)\bigr) \asymp M^2(M^2n)^{-4s/(4s + d)}~~\textrm{for $1 \leq d < 4s$,}
 \end{equation}
@@
 When $4s \geq d$ the functions in $H^s(\Xset)$ are very irregular---formally speaking $H^s(\Xset)$ does not continuously embed into $\Leb^4(\Xset)$ when $4s \geq d$---and as far as we know, the minimax testing rates in this regime have not been worked out.
-However, if one explicitly assumes that $f_0 \in \Leb^4(\mc{X})$ then the critical radius is given by $\epsilon^2 \asymp Mn^{-1/2}$.
+However, if one explicitly assumes that $f_0 \in \Leb^4(\mc{X})$ then the critical radius is given by $\epsilon_n^2(\Leb^4(\mc{X};M)) \asymp Mn^{-1/2}$.
 As a sanity check note that this is strictly worse than the rate in~\eqref{eqn:sobolev_space_testing_critical_radius}.
@@
 Since $\wh{f}$ is defined only at the design points $X_1,\ldots,X_n$, we will measure estimation rates in $\|\cdot\|_n^2$ error.
-The lower bounds used to show~\eqref{eqn:sobolev_space_minimax_estimation_rate} will still hold when in-sample loss is used rather than out-of-sample loss \textcolor{red}{(?)}, and we verify matching upper bounds in the sections to come.
+The lower bounds used to show~\eqref{eqn:sobolev_space_minimax_estimation_rate} will still hold when in-sample loss is used rather than out-of-sample loss \textcolor{red}{(TODO): (I don't actually know if this is true.
+Need to check.)}, and we verify matching upper bounds in the sections to come.
 Thus the minimax rates are unchanged---we call estimators which achieve these optimal rates \emph{in-sample minimax rate-optimal estimators}.
@@
 There is one other subtlety introduced by the use of in-sample mean squared error.
-Technically speaking, elements $f \in H^s(P)$ are equivalence classes, defined only up to a set of measure zero.
+Technically speaking, elements $f \in H^s(\mc{X})$ are equivalence classes, defined only up to a set of measure zero.
 Thus one cannot speak of the pointwise evaluation $f_0(X_i)$, as we do by defining our target of estimation to be $f_0(X_i)$, $i=1,\ldots,n$, until one selects \emph{representatives}.
-When $s > d/2$, every element $f$ of $H^s(P)$ admits a continuous version $f^{\ast}$, and it is standard to pick this as the favored representative.
-When $s \leq d/2$, some elements in $H^s(P)$ do not have any continuous version; however they admit a \emph{quasi-continuous} version \citep{evans15} known as the \emph{precise representative}, and we use this representative.
+When $s > d/2$, every element $f$ of $H^s(\mc{X})$ admits a continuous version $f^{\ast}$, and it is standard to pick this as the favored representative.
+When $s \leq d/2$, some elements in $H^s(\mc{X})$ do not have any continuous version; however they admit a \emph{quasi-continuous} version \citep{evans15} known as the \emph{precise representative}, and we use this representative.
 To be clear, however, it does not really matter which representative we choose.
@@
 As previously explained, Laplacian eigenmaps is a discrete and noisy approximation to a spectral series method, of particular use when the design distribution $P$ is unknown.
-Our goal is to show that Laplacian eigenmaps are rate optimal methods, notwithstanding the extra error incurred by this approximation.
+Our goal is to show that the Laplacian eigenmaps methods are rate optimal, notwithstanding the extra error incurred by this approximation.
 In this section and the following one, we will see that this is indeed the case: the estimator $\wh{f}$, and a test using the statistic $\wh{T}$, achieve optimal estimation and goodness-of-fit testing rates over Sobolev classes.
@@
 \label{sec:first_order_sobolev_classes}
-We begin by assuming $f_0 \in H^1(P)$.
-We show that $\wh{f}$ and a test based on $\wh{T}$ are minimax optimal, for all values of $d$, and under no additional assumptions on the data generating process, i.e.
+We begin by assuming $f_0 \in H^1(\mc{X})$.
+We show that $\wh{f}$ and a test based on $\wh{T}$ are minimax optimal, for all values of $d$, and under no additional assumptions (beyond those of Model~\ref{def:model_flat_euclidean}) on the data generating process, i.e.
 on either $P$ or $f_0$.
 
-\paragraph{Estimation.} When the graph radius $\varepsilon$ and number of eigenvectors $K$ are chosen appropriately, we show in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that the estimator $\wh{f}$ achieves the minimax rate over $H^1(P;M)$, when error is measured in $\|\cdot\|_{n}$.
+\paragraph{Estimation.} When the kernel $\eta$, graph radius $\varepsilon$, and number of eigenvectors $K$ are chosen appropriately, we show in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that the estimator $\wh{f}$ achieves the minimax rate over $H^1(\mc{X};M)$, when error is measured in $\|\cdot\|_{n}$.
 \begin{enumerate}[label=(A\arabic*)]
 	\setcounter{enumi}{0}
+	\item
+	\label{asmp:kernel_flat_euclidean}
+	The kernel function $\eta$ is a nonincreasing function supported on $[0,1]$. Its restriction to $[0,1]$ is Lipschitz, and $\eta(1) > 0$. Additionally, it is normalized so that
+	\begin{equation*}
+	\int_{\Rd} \eta(\|z\|) \,dz = 1.
+	\end{equation*}
+	and we assume \smash{$\int_{\Rd} \|x\|^2 \eta(\|x\|) \,dx < \infty$}.
 	\item 
@@
 	\label{thm:laplacian_eigenmaps_estimation_fo}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H^1(P,M)$.
-There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_fo}, then
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H^1(\mc{X},M)$.
+There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_fo}, then
 	\begin{equation}
@@
 \begin{enumerate}[label=(A\arabic*)]
-	\setcounter{enumi}{1}
+	\setcounter{enumi}{2}
 	\item 
@@
 Then $\mathbb{E}_0[\varphi] \leq a$, i.e $\varphi$ is a level-$a$ test.
-Suppose additionally $f_0 \in H^1(P,M)$, and that $d < 4$.
-Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_testing_fo}, and if $f_0$ satisfies
+Suppose additionally $f_0 \in H^1(\mc{X},M)$, and that $d < 4$.
+Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_fo}, and if $f_0$ satisfies
 	\begin{equation}
@@
 \end{theorem}
-Although~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} involves taking the maximum of several different terms, the important takeaway of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} is that if $n^{-1/2} \lesssim M^2 \ll n^{1/4}$ ($d = 1$) or $n^{-1/2} \lesssim M^2 \ll n^{2/d - 1/2}$ ($d = 2$ or $d = 3$), then $\varphi$ is a minimax rate-optimal test over $H^1(P,M)$.
+Although~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} involves taking the maximum of several different terms, the important takeaway of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} is that if $n^{-1/2} \lesssim M^2 \ll n^{1/4}$ ($d = 1$) or $n^{-1/2} \lesssim M^2 \ll n^{2/d - 1/2}$ ($d = 2$ or $d = 3$), then $\varphi$ is a minimax rate-optimal test over $H^1(\mc{X},M)$.
 
@@
 Indeed, as far we know when $d \geq 4$ the optimal testing rates over $H^1(\mc{X})$ are not known.
-On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then we show in Appendix~\textcolor{red}{(?)} that the Laplacian eigenmaps test, with $K = n$, is minimax optimal whenever $\|f_0\|_n^2 \gtrsim n^{-1/2}$.
-	\item The requirement that $\|f_0\|_P^2 \gtrsim n^{-p(d)}$ is necessary in order for us to establish that with probability at least $1 - 5\delta$,
+On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then we show in Appendix~\textcolor{red}{(?)} that the Laplacian eigenmaps test, with $K = n$, has small type II error whenever $\|f_0\|_P^2 \gtrsim n^{-1/2}$.
+	\item The requirement in Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} that $\|f_0\|_P^2 \gtrsim n^{-p(d)}$ is necessary in order for us to establish that with probability at least $1 - 5\delta$,
 	\begin{equation}
@@
 We now assume that the regression function displays some higher-order regularity and is zero-trace, i.e.
-$f_0 \in H_0^s(P)$.
+$f_0 \in H_0^s(\mc{X})$.
 At first glance, it is not at all clear that Laplacian eigenmaps \emph{can} take advantage of this additional regularity, and it is especially unclear whether it can do so in an optimal manner.
@@
 \paragraph{Estimation.}
-In order to show that $\wh{f}$ is an optimal estimator over $H_0^s(P;M)$, we will require that $\varepsilon$ be meaningfully larger than the lower bound in~\ref{asmp:parameters_estimation_fo}.
+In order to show that $\wh{f}$ is an optimal estimator over $H_0^s(\mc{X};M)$, we will require that $\varepsilon$ be meaningfully larger than the lower bound in~\ref{asmp:parameters_estimation_fo}.
 \begin{enumerate}[label=(A\arabic*)]
-	\setcounter{enumi}{2}
+	\setcounter{enumi}{3}
 	\item 
@@
 	\label{thm:laplacian_eigenmaps_estimation_ho}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H_0^s(P,M)$ and $p \in C^{s - 1}(\mc{X})$.
-There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_ho}, then
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H_0^s(\mc{X},M)$ and $p \in C^{s - 1}(\mc{X})$.
+There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_ho}, then
 	\begin{equation}
@@
 \end{theorem}
-Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that in the flat Euclidean setting Laplacian eigenmaps is a rate-optimal estimator over Sobolev classes, for all values of $s$ and $d$.
+Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that in the flat Euclidean setting Laplacian eigenmaps is an in-sample minimax rate optimal estimator over Sobolev classes, for all values of $s$ and $d$.
 Some other remarks:
@@
 Let $\wc{f}$ be the Laplacian eigenmaps estimator computed using the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ .
-Because $\wc{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wc{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$, and consequently much faster to compute $\wc{f}$ than $\wh{f}$\textcolor{red}{(reference needed)}.
+Because $\wc{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wc{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$, and consequently much faster to compute $\wc{f}$ than $\wh{f}$ \textcolor{red}{(reference needed)}.
 Statistically speaking, as we show in Appendix~\textcolor{red}{(?)} $\wc{f}$ has $L^2(P_n)$-error of at most $\sigma^{2s}$ times the error of $\wh{f}$.
@@
 \begin{enumerate}[label=(A\arabic*)]
-	\setcounter{enumi}{3}
+	\setcounter{enumi}{4}
 	\item 
@@
 Then $\mathbb{E}_0[\varphi] \leq a$, i.e $\varphi$ is a level-$a$ test.
-Suppose additionally $f_0 \in H_0^s(P,M)$, that $p \in C^{s-1}(\mc{X})$, and that $d \leq 4$.
-Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_testing_ho}, and if $f_0$ satisfies
+Suppose additionally $f_0 \in H_0^s(\mc{X},M)$, that $p \in C^{s-1}(\mc{X})$, and that $d \leq 4$.
+Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_ho}, and if $f_0$ satisfies
 	\begin{equation}
@@
 \end{theorem}
-Similarly to the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that $\varphi$ is a minimax optimal test over $H_0^s(P)$ when $n^{-1/2} \ll M^2 \ll n^{1/4}$ (if $2s < d$), or if $n^{-1/2} \ll M^2 \ll (2/d - 1/2)$ (if $2s \geq d$.)
-
-However, unlike the first-order case, when $4 < d < 4s$ the minimax testing rate over $H_0^s(P)$ is still $n^{-4s/(4s + d)}$.
+Similarly to the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that $\varphi$ is a minimax optimal test over $H_0^s(\mc{X})$ when $n^{-1/2} \ll M^2 \ll n^{1/4}$ (if $2s < d$), or if $n^{-1/2} \ll M^2 \ll (2/d - 1/2)$ (if $2s \geq d$.)
+
+However, unlike the first-order case, when $4 < d < 4s$ the minimax testing rate over $H_0^s(\mc{X})$ is still $n^{-4s/(4s + d)}$.
 Unfortunately, we can no longer claim that $\varphi$ is an optimal test in this regime.
-There do not exist any choices of $\varepsilon$ and $K$ which satisfy both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho}.
-As a result we can't balance testing bias and variance (defined momentarily in~\eqref{eqn:testing_biasvariance}).
-It follows from Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} and the triangle inequality that, choosing $K \asymp (n)^{-2s/(2s + d)}$ as is optimal for estimation, $\varphi$ must have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2s + d)}$.
+There do not exist any choices of $\varepsilon$ and $K$ which satisfy both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho}, and as a result we cannot optimally balance testing bias and variance (defined momentarily in~\eqref{eqn:testing_biasvariance}).
+It follows from Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} that, choosing $K \asymp (n)^{-2s/(2s + d)}$ as is optimal for estimation, $\varphi$ must have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2s + d)}$.
 In fact, we can derive slightly better (though still not optimal) rates.
@@
 	Under the same setup as Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, but with $d > 4$.
-If the Laplacian eigenmaps test $\varphi$ is computed with $K$ satisfying~\eqref{eqn:eigenvector_testing_ho}, and $\varepsilon \asymp K^{-1/d}$, and if
+If the Laplacian eigenmaps test $\varphi$ is computed kernel $\eta$ satisfying~\eqref{asmp:kernel}, $K$ satisfying~\eqref{eqn:eigenvector_testing_ho}, and $\varepsilon \asymp K^{-1/d}$, and if
 	\begin{equation}
@@
 
+\textcolor{red}{(TODO): This last bit is too vague and speculative. Maybe worth deleting entirely?}
+
 A last point: as mentioned at the beginning of this section, the estimation and testing rates we obtain in Theorems~\ref{thm:laplacian_eigenmaps_estimation_ho}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} are superior to those implied by the concentration of graph Laplacian eigenvectors $v_k$ around eigenfunctions $\psi_k$ of the weighted Laplace-Beltrami operator $\Delta_P$.
@@
 \begin{enumerate}
-	\item The convergence rates of $v_k \to \psi_k$ are faster than currently known,  particularly when $p$ exhibits higher-order smoothness.
+	\item The convergence rates of graph eigenvectors to continuum eigenfunctions, $v_k \to \psi_k$, are faster than currently known,  particularly if $p$ exhibits higher-order smoothness.
 	\item The approximation properties of $v_1,\ldots,v_K$ are much stronger than is implied by the concentration of each individual eigenvector $v_k$ around its continuum counterpart $\psi_k$.
@@
 Identifying which explanation is responsible would, in our view, be a useful step towards understanding the success of graph Laplacian methods.
-\textcolor{red}{(TODO): Too vague and speculative.
-Maybe worth deleting entirely?}
 
@@
 \end{equation}
-The graph quadratic form $\dotp{L^s f_0}{f_0}_n$, eigenvalue $\lambda_{K}$, and empirical squared norm $\|f_0\|_n^2$ are each random variables that depend the random design points $X_1,\ldots,X_n$.
-We now establish suitable upper and lower bounds on these quantities.
+The quadratic form $\dotp{L_{n,\varepsilon}^s f_0}{f_0}_n$, eigenvalue $\lambda_{K}$, and empirical squared norm $\|f_0\|_n^2$ are each random variables that depend the random design points $X_1,\ldots,X_n$.
+We proceed to establish suitable upper and lower bounds on these quantities.
 
 \paragraph{Estimates on graph quadratic forms.}
-We begin by restating in Proposition~\ref{prop:graph_seminorm_fo} an upper bound on the first-order quadratic form $\dotp{L f_0}{f_0}_n$ from \textcolor{blue}{Green et al. 2021}.
+We begin by restating in Proposition~\ref{prop:graph_seminorm_fo} an upper bound on the first-order quadratic form $\dotp{L_{n,\varepsilon}f_0}{f_0}_n$ from \textcolor{blue}{Green et al. 2021}.
 \begin{proposition}[Lemma~1 of \textcolor{blue}{Green et al. 2021}]
 	\label{prop:graph_seminorm_fo}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H^1(P,M)$.
-There exist constants $c,C$ that do not depend on $f_0$ such that the following statement holds for any $\delta \in (0,1)$: if $\varepsilon < c$, then
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H^1(\mc{X},M)$.
+There exist constants $c,C$ that do not depend on $f_0$ such that the following statement holds for any $\delta \in (0,1)$: if $\eta$ satisfies~\ref{asmp:kernel} and $\varepsilon < c$, then
 	\begin{equation}
@@
 	\label{prop:graph_seminorm_ho} 
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f_0 \in H_0^s(P;M)$ and $p \in C^{s - 1}(\mc{X})$.
-Then there exist constants $c$ and $C$ that do not depend on $f_0$ such that the following statement holds for any $\delta \in (0,1)$: if $Cn^{-1/(2(s - 1) + d)} < \varepsilon < c$, then
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f_0 \in H_0^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$.
+Then there exist constants $c$ and $C$ that do not depend on $f_0$ such that the following statement holds for any $\delta \in (0,1)$: if $\eta$ satisfies~\ref{asmp:kernel} and $Cn^{-1/(2(s - 1) + d)} < \varepsilon < c$, then
 	\begin{equation}
@@
 
-The analysis of $\dotp{L_{n,\varepsilon}^sf_0}{f_0}$ when $s > 1$ is substantially trickier than the first-order case, for several reasons.
-First of all, we note that when $s > 1$ the graph energy $\dotp{L_{n,\varepsilon}^s f}{f}_n$ is itself a biased estimate of the non-local energy $E_{\varepsilon}^{(s)}(f)$, in contrast to the first order case where $\Ebb\bigl[\dotp{L_{n,\varepsilon} f}{f}_n\bigr] = E_{\varepsilon}(f)$.
-In the language of classical statistics, $\dotp{L_{n,\varepsilon}^s f}{f}_n$ is a $V$-statistic, that is the sum of an unbiased estimator of $E_{\varepsilon}^{(s)}(f)$ (in other words, a $U$-statistic) plus some higher-order, pure bias terms.
+There are several reasons why this analysis is substantially trickier than the first-order case.
+First of all, we note that when $s > 1$ the quadratic form $\dotp{L_{n,\varepsilon}^s f}{f}_n$ is itself a biased estimate of the non-local energy $E_{\varepsilon}^{(s)}(f)$, in contrast to the first order case where $\Ebb\bigl[\dotp{L_{n,\varepsilon} f}{f}_n\bigr] = E_{\varepsilon}^{(1)}(f)$.
+One can view $\dotp{L_{n,\varepsilon}^s f}{f}_n$ as a $V$-statistic, meaning it is the sum of an unbiased estimator of $E_{\varepsilon}^{(s)}(f)$ (in other words, a $U$-statistic) plus some higher-order, pure bias terms.
 These pure bias terms are negligible only when $\varepsilon \gg n^{-1/(2r + d)}$.
 
-On the other hand, in order to give an estimate of $E_{\varepsilon}^{(s)}(f)$ in terms of the integrated, squared derivatives of $f_0$ and $p$, we first show that $L_{P,\varepsilon}^jf(x) \to \Delta_P^jf(x)$ as $\varepsilon \to 0$; here $j = (s - 1)/2$ when $s$ is odd and $j = (s - 2)/2$ when $s$ is even.
+On the other hand, in order to give an estimate of $E_{\varepsilon}^{(s)}(f)$ in terms of the Sobolev norm of $f_0$ and H\"{o}lder norm of $p$, we first show that $L_{P,\varepsilon}^jf(x) \to \Delta_P^jf(x)$ as $\varepsilon \to 0$; here $j = (s - 1)/2$ when $s$ is odd and $j = (s - 2)/2$ when $s$ is even.
 In words, we show that a $j$th-order iterated difference operator approximates a $2j$-th order differential operator.
 In order to establish this fact, we require that on the one hand that $p \in C^{s - 1}(\mc{X})$, and on the other hand that $f$ be zero-trace.
-Crucially we do not try to show $L_{P,\varepsilon}^sf(x) \to \Delta_{P}^sf(x)$: though this may seem like a natural first step, note carefully that the latter object is an order-$2s$ differential operator, whereas we assume that $f$ has only $s$ bounded derivatives.
+Crucially we do not try to show $L_{P,\varepsilon}^sf(x) \to \Delta_{P}^sf(x)$: though this may seem like a natural first step towards establishing $\dotp{L_{P,\varepsilon}^sf}{f}_{P} \to \dotp{\Delta_P^sf}{f}_{P}$, note carefully that $\Delta_P^s$ is an order-$2s$ differential operator, whereas we assume that $f$ has only $s$ bounded derivatives.
 
@@
 
-In the case where $\mc{X}$ is assumed to have a boundary, the graph Laplacian $L$ is a reasonable approximation of the operator $\Delta_P$ only at interior points $x$ such that $B(x,\varepsilon) \subseteq \mc{X}$.
-In contrast, at points $x$ near the boundary $L_{n,\varepsilon}$ is known to approximate a completely different operator altogether \citep{belkin2012}.
+In the case where $\mc{X}$ is assumed to have a boundary, the graph Laplacian $L_{n,\varepsilon}$ is a reasonable approximation of the operator $\Delta_P$ at points $x \in \mc{X}$ for which $B(x,\varepsilon) \subseteq \mc{X}$.
+In contrast, at points $x$ near the boundary of $\mc{X}$, the graph Laplacian is known to approximate a different operator altogether \citep{belkin2012}.
 This is reminiscent of the boundary effects present in the analysis of kernel smoothing.
-Thus proving convergence of $\lambda_k$ to a continuum limit becomes a substantially more challenging problem.
-We will instead use the results of \textcolor{red}{(Green et al.
+Thus proving convergence of $\lambda_k$ to a continuum limit becomes a substantially more challenging problem when $\mc{X}$ has a boundary.
+Rather than establishing such a result, we will instead directly use Lemma~2 of \textcolor{red}{(Green et al.
 2021)}, whose assumptions match our own, and who give a weaker bound on $\lambda_k/\lambda_k(\Delta_P)$ that will nevertheless suffice for our purposes.
@@
 	\label{prop:graph_eigenvalue}
-	Suppose Model~\ref{def:model_flat_euclidean}. Then there exist constants $c$ and $C$ such that the following statement holds: if $C(\log n/n)^{1/d} < \varepsilon < c$, then
+	Suppose Model~\ref{def:model_flat_euclidean}.
+Then there exist constants $c$ and $C$ such that the following statement holds: if $\eta$ satisfies~\ref{asmp:kernel} and $C(\log n/n)^{1/d} < \varepsilon < c$, then
 	\begin{equation}
@@
 \paragraph{Empirical norm.}
-\textcolor{red}{(TODO)}
-
-Invoking the bounds of Propositions~\ref{prop:graph_seminorm_fo}-\ref{prop:graph_eigenvalue} inside the bias-variance tradeoffs~\eqref{eqn:estimation_biasvariance} and~\eqref{eqn:testing_biasvariance} and then choosing $K$ to balance bias and variance (if possible), leads to the conclusions of Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}.
+Finally, in order to show that $\varphi$ has small Type II error whenever~$\|f_0\|_P$ is greater than the critical radius given by~\eqref{eqn:sobolev_space_testing_critical_radius}, we require a lower bound on $\|f_0\|_n^2$ in terms of $\|f_0\|_P^2$.
+In Lemma~\ref{lem:empirical_norm_sobolev} we establish that such a bound holds, whenever $\|f_0\|_P$ is sufficiently large.
+When $f_0 \in H_0^s(\mc{X})$ and $s > d/2$ this is a standard estimate \textcolor{red}{(reference)}, but it seems to be less well-known in the subcritical regime $s \leq d/2$.
+\begin{proposition}
+	\label{prop:empirical_norm_sobolev}
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f_0 \in H^s(\Xset,M)$.
+There exists a constant $C$ that does not depend on $f_0$ such that the following statement holds for all $\delta \in (0,1)$:  if
+	\begin{equation}
+	\label{eqn:empirical_norm_sobolev_1}
+	\norm{f_0}_{P} \geq \frac{C M}{\delta}\Bigl(n^{-1/2} \vee n^{-s/d}\Bigr)
+	\end{equation}
+	then
+	\begin{equation}
+	\label{eqn:empirical_norm_sobolev}
+	\norm{f_0}_n^2 \geq \delta \|f_0\|_P^2
+	\end{equation}
+	with probability at least $1 - 5\delta$.
+\end{proposition}
+
+Finally, invoking the bounds of Propositions~\ref{prop:graph_seminorm_fo}-\ref{prop:empirical_norm_sobolev} inside the bias-variance tradeoffs~\eqref{eqn:estimation_biasvariance} and~\eqref{eqn:testing_biasvariance} and then choosing $K$ to balance bias and variance (when possible), leads to the conclusions of Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}.
 
@@
 In this section we consider the manifold setting, where $(X_1,Y_1),\ldots,(X_n,Y_n)$ are observed according to Model~\ref{def:model_manifold}.
-\citet{belkin03,belkin05,niyogi2013} establish that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
-We build on this work by showing that when $f_0 \in H^s(P)$ and $P$ is supported on a manifold, Laplacian eigenmaps achieve the sharper minimax estimation and testing rates reviewed in Section~\ref{subsec:minimax_rates_sobolev}.
-
-\subsection{Error rates under the manifold hypothesis}
-We now show that Laplacian eigenmaps achieves the faster minimax rates under the manifold hypothesis, when the regression function $f_0 \in H^s(P;M)$, $s \leq 4$, and, in the testing problem, $m \leq 4$.
-This section will proceed in a similar fashion to Section~\ref{sec:higher_order_sobolev_classes}, except with the ambient dimension $d$ replaced by intrinsic dimension $m$.
+A theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+We build on this work by showing that when $f_0 \in H^s(\mc{X})$ and $P$ is supported on a manifold, Laplacian eigenmaps achieve the sharper minimax estimation and testing rates reviewed in Section~\ref{subsec:minimax_rates_sobolev}.
+
+\subsection{Laplacian eigenmaps error rates under the manifold hypothesis}
+Unlike in the flat-Euclidean case, because in Model~\ref{def:model_manifold} we assume that $\mc{X}$ is boundaryless it is easy to deal with the first-order $(s = 1)$ and higher-order $(s > 1)$ cases all at once.
+A more distinction between the results we obtain in this section and those of the last is that we will establish Laplacian eigenmaps is optimal only when the regression function $f_0 \in H^s(\mc{X};M)$ for $s \leq 3$.
+Otherwise, this section will proceed in a similar fashion to Section~\ref{sec:higher_order_sobolev_classes}.
 
 \paragraph{Estimation.}
-To ensure that $\wh{f}$ is an in-sample minimax rate-optimal estimator, we choose the graph radius $\varepsilon$ and number of eigenvectors $K$ as in~\ref{asmp:parameters_estimation_ho}, except with ambient dimension $d$ replaced by the intrinsic dimension $m$.
+To ensure that $\wh{f}$ is an in-sample minimax rate-optimal estimator, we choose the kernel function $\eta$, graph radius $\varepsilon$ and number of eigenvectors $K$ as in~\ref{asmp:parameters_estimation_ho}, except with ambient dimension $d$ replaced by the intrinsic dimension $m$.
 
@@
 	\item 
+	\label{asmp:kernel_manifold}
+	The kernel function $\eta$ is a nonincreasing function supported on a subset of $[0,1]$. Its restriction to $[0,1]$ is Lipschitz, and $\eta(1/2) > 0$. Additionally, it is normalized so that
+	\begin{equation*}
+	\int_{\Reals^m} \eta(\|z\|) \,dz = 1,
+	\end{equation*}
+	and we assume \smash{$\int_{\Reals^m} \|x\|^2 \eta(\|x\|) \,dx < \infty$}.
+	\item 
 	\label{asmp:parameters_estimation_manifold}
@@
 	\label{thm:laplacian_eigenmaps_estimation_manifold}
-	Suppose Model~\ref{def:model_manifold}, and additionally $f_0 \in H^s(P,M)$ and $p \in C^{s - 1}(\mc{X})$ for $s < 4$.
-There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_manifold}, then
+	Suppose Model~\ref{def:model_manifold}, and additionally $f_0 \in H^s(\mc{X},M)$ and $p \in C^{s - 1}(\mc{X})$ for $s \leq 3$.
+There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_manifold}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_manifold}, then
 	\begin{equation}
@@
 \begin{enumerate}[label=(A\arabic*)]
-	\setcounter{enumi}{5}
+	\setcounter{enumi}{7}
 	\item 
@@
 Then $\mathbb{E}_0[\varphi] \leq a$, i.e $\varphi$ is a level-$a$ test.
-Suppose additionally $f_0 \in H^s(P,M)$, that $p \in C^{s-1}(\mc{X})$, and that $s < 4$ and $m \leq 4$.
-Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_testing_manifold}, and if $f_0$ satisfies
+Suppose additionally $f_0 \in H^s(\mc{X},M)$, that $p \in C^{s-1}(\mc{X})$, and that $s \leq 3$ and $m \leq 4$.
+Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed kernel $\eta$ satisfying~\ref{asmp:kernel_manifold}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_manifold}, and if $f_0$ satisfies
 	\begin{equation}
@@
 The difference is that when $\mc{X}$ is a manifold with intrinsic dimension $m$, we can prove analogous results to Propositions~\ref{prop:graph_seminorm_fo}-\ref{prop:graph_eigenvalue}, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
-	\item Unlike in the full-dimensional case, our upper bounds on the estimation and testing error of Laplacian eigenmaps match the minimax rate only when $s < 4$.
-Our upper bounds when $s \geq 4$ follow from the embedding $H^s(P;M) \subset H^{3}(P;M)$, i.e they match the rates we get just by assuming that the $3$rd order derivative is bounded, and are clearly suboptimal.
+	\item Unlike in the full-dimensional case, our upper bounds on the estimation and testing error of Laplacian eigenmaps match the minimax rate only when $s \leq 3$.
+Our upper bounds when $s \geq 4$ follow from the embedding $H^s(\mc{X};M) \subset H^{3}(P;M)$, i.e they match the rates we get just by assuming that the $3$rd order derivative is bounded, and are clearly suboptimal.
 
@@
 
-Of course, the Laplacian eigenmaps estimator is only defined at the observed design points $X_1,\ldots,X_n$, and to measure its error in $L^2(P)$ norm we must first extend it to the rest of $\Xset$.
+Of course, the Laplacian eigenmaps estimator is only defined at the observed design points $X_1,\ldots,X_n$, and to measure its error in $L^2(P)$ norm we must first extend it to be defined over all of $\Xset$.
 We propose a simple method, kernel smoothing, to do the job.
-The method can applied to any estimator defined at the design points, including Laplacian eigenmaps, and we show that a smoothed version $\wt{f}$ of our original estimator $\wh{f}$ has optimal $L^2(P)$ error.
-For simplicity, in this section we stick to the full-dimensional setting, but everything can be adapted to the manifold setting in a straightforward way.
+The method can applied to any estimator defined at the design points, including Laplacian eigenmaps, and we show that a smoothed version of our original estimator $\wh{f}$ has optimal $L^2(P)$ error.
+For simplicity, in this section we will stick to the flat Euclidean setting, where $(X_1,Y_1),\ldots,(X_n,Y_n)$ are observed according to Model~\ref{def:model_flat_euclidean}.
 
@@
 We now formally define our approach to extension by kernel smoothing.
-For a kernel function $\psi(\cdot): [0,\infty) \to (-\infty,+\infty)$, bandwidth $h > 0$, and a distribution $Q$, the \emph{Nadaraya-Watson kernel smoother} $T_{\varepsilon,Q}$ is given by
-\begin{equation*}
-\bigl(T_{h,Q}f)(x) := \frac{1}{d_Q(x)} \int_{\Omega} f(z)\psi\biggl(\frac{\|z - x\|}{h}\biggr) \,dQ(z)
-\end{equation*}
-where $d_Q(x) := \int_{\Omega} \psi\bigl(\|z - x\|/\varepsilon\bigr) \,dQ(z)$. We denote $T_{\varepsilon,P_n}$ by $T_{\varepsilon,n}$, and $d_{P_n}$ by $d_n$.
-
-The estimated function $\wt{f} \in L^2(P)$ we will consider is a kernel smoother passed over $\wh{f}$, meaning
-\begin{equation}
-\label{eqn:kernel_smoother_laplacian_eigenmaps}
-\wt{f}(x) := \bigl(T_{h,n}\wh{f}\bigr)(x).
-\end{equation}
-It is useful to make a few comments about $\wt{f}$.
-First of all, while in principle~\eqref{eqn:kernel_smoother_laplacian_eigenmaps} makes sense at any point $x \in \Rd$, we will always think of $\wt{f}$ as a function mapping $\mc{X}$ to $\Reals$.
-Additionally, note we may use the same kernel for smoothing as was used to construct the graph, i.e.
-take $\psi = \eta$.
-Likewise, we can set $h = \varepsilon$.
-However, we will see that for theoretical purposes these are not always the best choices.
-
-\paragraph{Out-of-sample error of kernel smoothed Laplacian Eigenmaps.}
+For a kernel function $\psi(\cdot): [0,\infty) \to (-\infty,+\infty)$, bandwidth $h > 0$, and a distribution $Q$, the \emph{Nadaraya-Watson kernel smoother} $T_{h,Q}$ is given by
+\begin{equation*}
+\bigl(T_{h,Q}f)(x) := \frac{1}{d_Q(x)} \int_{\Omega} f(z)\psi\biggl(\frac{\|z - x\|}{h}\biggr) \,dQ(z),
+\end{equation*}
+where $d_Q(x) := \int_{\Omega} \psi\bigl(\|z - x\|/\varepsilon\bigr) \,dQ(z)$.
+For convenience, we will denote $T_{\varepsilon,P_n}$ by $T_{\varepsilon,n}$, and $d_{P_n}$ by $d_n$.
+We extend the Laplacian eigenmaps estimator by passing the kernel smoother $T_{h,n}$ over it, that is we consider the estimator $T_{h,n}\wh{f}$, which is defined at every $x \in \mc{X}$.
+
+\paragraph{Out-of-sample error of kernel smoothed Laplacian eigenmaps.}
 
 In Lemma~\ref{lem:kernel_smoothing_laplacian_eigenmaps}, we consider an arbitrary estimator $\wc{f} \in L^2(P_n)$.
-We show that the out-of-sample error $\|T_{n,h}\wc{f} - f_0\|_P^2$ can be upper bounded by three terms--- (a constant times) the in-sample error of $\wc{f} - f_0$, and variance and bias terms associated with a noiseless version of kernel smoothing.
+We show that the out-of-sample error $\|T_{n,h}\wc{f} - f_0\|_P^2$ can be upper bounded by three terms--- (a constant times) the in-sample error $\|\wc{f} - f_0\|_n^2$, and variance and bias terms that arise naturally in the analysis of kernel smoothing over noiseless data..
 We shall assume the following conditions on $\psi$.
@@
 	\label{lem:kernel_smoothing_laplacian_eigenmaps}
-	Suppose $\wc{f} \in L^2(P_n)$, $f_0 \in H^1(P)$ and $p \in C^1(P)$. Under Assumptions~\ref{asmp:domain}, \ref{asmp:density}, and~\ref{asmp:kernel}, it holds that
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $\wc{f} \in L^2(P_n)$, $f_0 \in H^1(\mc{X})$ and $p \in C^1(\mc{X})$.
+If the kernel smoothing estimator $T_{h,n}\wc{f}$ is computed with kernel $\psi$ satisfying~\ref{asmp:kernel}, it holds that
 	\begin{equation}
@@
 \begin{equation*}
-\int_{-\infty}^{\infty} \psi(|s|) = 1,~~\int_{-\infty}^{\infty} t^j \psi(|t|) \,dt = 0~~\textrm{for $j = 1,\ldots,s - 1$},~~\textrm{and}~~\int_{-\infty}^{\infty} t^s \psi(|t|) \,dt < \infty.
-\end{equation*}
+\int_{-\infty}^{\infty} \psi(|s|) = 1,~~\int_{-\infty}^{\infty} t^j \psi(|t|) \,dt = 0~~\textrm{for $j = 1,\ldots,s + d - 2$},~~\textrm{and}~~\int_{-\infty}^{\infty} t^{s + d - 1} \psi(|t|) \,dt < \infty.
+\end{equation*}
+
+\textcolor{red}{(TODO): I have worked out the math for this result, but I am a little surprised at the requirement that $\psi$ need annihilate polynomials up to degree $s + d - 2$.
+Will need to double check closely, since I could not find a reference analyzing kernel smoothing using a radially symmetric higher order kernel when $d > 1$.}
 \begin{lemma}
 	\label{lem:kernel_smoothing_bias}
-	Suppose the domain $\mc{X}$, the distribution $P$, and the kernel $\psi$  satisfy~\ref{asmp:domain}, \ref{asmp:density}, and~\ref{asmp:kernel}, respectively. 
+	Suppose Model~\ref{def:model_flat_euclidean}, and that the smoothing operator $T_{h,P}$ is computed with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel}.
 	\begin{itemize}
-		\item If $f_0 \in H^1(P)$, it follows that
-		\begin{equation*}
-		\|T_{h,P}f_0 - f_0\|_P^2 \leq C h^{2} |f|_{H^1(P)}^2.
+		\item If $f_0 \in H^1(\mc{X})$, it follows that
+		\begin{equation*}
+		\|T_{h,P}f_0 - f_0\|_P^2 \leq C h^{2} |f|_{H^1(\mc{X})}^2.
 		\end{equation*}
@@
 		\begin{equation*}
-		\|T_{h,P}f_0 - f_0\|_P^2 \leq C h^{2s} |f|_{H^s(P)}^2.
+		\|T_{h,P}f_0 - f_0\|_P^2 \leq C h^{2s} |f|_{H^s(\mc{X})}^2.
 		\end{equation*}
@@
 
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_estimation_out_of_sample}
-	Suppose we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to~\eqref{eqn:model}.
-Suppose that for some $s \in \mathbb{N}$, the regression function $f_0 \in H^s(P;M)$ (if $s = 1$) or $H_0^s(P;M)$ (if $s > 1$).
+% AG: Get rid of this!
+% \begin{theorem}
+%	\label{thm:laplacian_eigenmaps_estimation_out_of_sample}
+%	Suppose we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to~\eqref{eqn:model}.
+Suppose that for some $s \in \mathbb{N}$, the regression function $f_0 \in H^s(\mc{X};M)$ (if $s = 1$) or $H_0^s(\mc{X};M)$ (if $s > 1$).
 Suppose also that the distribution $P$ has support $\mc{X}$ which satisfies~\ref{asmp:domain} and density $p \in C^s(\mc{X};M)$ which satisfies~\ref{asmp:density}.
 Then if the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_ho}, then the out-of-sample extension $T_{h,n}\wh{f}$ computed with bandwidth $h \asymp n^{-1/(2(s - 1) + d)}$ satisfies
-	\begin{equation*}
-	\|T_{h,n}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^{2d/(2s + d)}n^{-2s/(2s + d)},
-	\end{equation*}
-	with probability at least $1 - \delta - Cn\exp\{-Cn\varepsilon^d\} - Ch^d\exp\{-Cnh^d\}$. 
-\end{theorem}
-Some remarks:
+%	\begin{equation*}
+%	\|T_{h,n}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^{2d/(2s + d)}n^{-2s/(2s + d)},
+%	\end{equation*}
+%	with probability at least $1 - \delta - Cn\exp\{-Cn\varepsilon^d\} - Ch^d\exp\{-Cnh^d\}$. 
+% \end{theorem}
+
+Some other remarks:
 \begin{itemize}
-	\item \textbf{Tuning parameters by sample-splitting.} Extending the estimator $\wh{f}$ to be defined out-of-sample allows us to select the hyper-parameters $\varepsilon$ and $K$ using sample splitting or cross validation.
+	\item Extending the estimator $\wh{f}$ to be defined out-of-sample allows us to select the hyper-parameters $\varepsilon$ and $K$ using sample splitting or cross validation.
 For instance, we can (i) split the sample into two halves, (ii) use the first half to compute $T_{h,n}\wh{f}$ for various values of $\varepsilon$, $h$, and $K$, (iii) choose the optimal values of these three hyperparameters by minimizing error on the held out set.
 This is by far the most common used way to choose hyperparameters \textcolor{red}{(references)}, and it addresses a serious limitation of Laplacian eigenmaps.
-Moreover, it is known \textcolor{red}{(references)} that choosing hyper-parameters by minimizing error over a held-out set results in estimators that optimally adapt to the order of regularity $s$---that is, that are rate-optimal (up to $\log$ factors), even when $s$ is unknown.
+Moreover, it is known \textcolor{red}{(references)} that choosing hyper-parameters by minimizing error over a held-out set can result in estimators that optimally adapt to the order of regularity $s$---that is, that are rate-optimal (up to $\log$ factors), even when $s$ is unknown.
 We believe that similar arguments will imply that $T_{h,n}\wh{f}$ is adaptive in this sense, though we do not pursue the details.
-	\item \textbf{Other extensions.} We consider extension by kernel smoothing because it is a simple and statistically optimal procedure that does not require any knowledge of the domain $\mc{X}$ or distribution $P$---as we have argued, this latter property is one of the main selling points of Laplacian eigenmaps as a tool for nonparametric regression.
-
-	That being said, there exist (many) other approaches to extending (or interpolating) a function $f$ based on the evaluations $\{f(X_1),\ldots,f(X_n)\}$, and we now comment on a few of these alternatives.
+\textcolor{red}{(TODO): Siva would like me to pursue the details.}
+	\item We consider extension by kernel smoothing because it is a simple and statistically optimal procedure that does not require any knowledge of the domain $\mc{X}$ or distribution $P$---as we have argued, this latter property is one of the main selling points of Laplacian eigenmaps as a tool for nonparametric regression.
+
+	That being said, there exist many other approaches to extending (or interpolating) a function $f$ using only evaluations $\{f(X_1),\ldots,f(X_n)\}$, and we now comment on a few of these alternatives.
 One approach is to perform minimum norm interpolation, where one defines a normed space $(\mc{F},\|\cdot\|_{\mc{F}})$ and then solves the optimization problem
@@
 Naturally, this approach works well when $f$ is close to a function $u \in \mc{F}$ with reasonably small norm.
-This holds true when $f \in H^s(P;M)$ and $s > d/2$, but as already discussed when $s \leq d/2$ the Sobolev space $H^s(P)$ is not an RKHS, and in fact when $s \leq d/2$ thin-plate spline interpolation is ill-posed \citep{green93}.
+This holds true when $f \in H^s(\mc{X};M)$ and $s > d/2$, but as already discussed when $s \leq d/2$ the Sobolev space $H^s(\mc{X})$ is not an RKHS, and in fact when $s \leq d/2$ thin-plate spline interpolation is ill-posed \citep{green93}.
 Another, arguably simpler approach is to extend $f$ to be piecewise constant on the Voronoi tessellation induced by $X_1,\ldots,X_n$, or equivalently to perform $1$-nearest neighbors regression on $f$.
-However, this approach is theoretically optimal only when $f \in C^1(P)$, in contrast to the kernel smoothing method we propose and study.
+However, this approach is theoretically optimal only when $f_0 \in C^1(\mc{X})$, in contrast to the kernel smoothing method we propose and study.
+
+	None of these three methods are intrinsically linked to Laplacian eigenmaps.
+This is in one sense a strength, since they can be used to extrapolate any estimator defined only in-sample.
+But it also potentially a weakness.
+Each of these methods have their own statistical and approximation errors (bias and variance) which can be fundamentally different than those of Laplacian eigenmaps, and there is a danger that in extrapolating the Laplacian eigenmaps estimator in this way we are taking the ``worst of both worlds''.
+Our theory shows that when the data model is Model~\ref{def:model_flat_euclidean} and we perform extrapolation by kernel smoothing, this is not a problem, at least in a minimax sense.
+
 	%Finally, the Nystr\"{o}m method extends  eigenvectors $v_k$ to functions $f_k: \mc{X} \to \Reals$ as follows,
@@
 
-\textcolor{red}{(TODO)}: Come up with a set of experiments that might be interesting to run.
-
+\textcolor{red}{(TODO): I list some potential options below.}
 \begin{itemize}
@@
 Investigate the MSE heat map of Laplacian eigenmaps, as a function of $K$ and $\varepsilon$.
-	\item Swiss roll, density \textcolor{red}{(?)}, regression function \textcolor{red}{(?)}. Evaluate Laplacian eigenmaps in-sample error, and Laplacian-eigenmaps + kernel smoothing out-of-sample error.
+	\item \textbf{Manifold setting.} Swiss roll, density \textcolor{red}{(?)}, regression function \textcolor{red}{(?)}.
+Evaluate Laplacian eigenmaps in-sample error, and Laplacian-eigenmaps + kernel smoothing out-of-sample error.
 \end{itemize}
@@
 The standard recommended alternative to spectral series methods, when the distribution $P$ is considered non-uniform or unknown, is to do least-squares.
-Concretely, suppose instead we had access only to eigenfunctions $f_1',f_2',\ldots$ of the unweighted Laplace-Beltrami operator $\Delta = \Delta_{\nu}$.
-Then letting $F_K = \mathrm{span}\{f_1,\ldots,f_K\}$, the least-squares estimator and test statistic
-\begin{equation*}
-\wt{f}_{\mathrm{LS}} := \argmin_{f \in F_K} \|Y - f\|_n^2,\textrm{and}~~\wt{T}_{\mathrm{LS}} = \|\wt{f}\|_{\nu}^2
-\end{equation*}
-are still rate-optimal over $H_0^s(P)$.
+Concretely, suppose instead we had access only to eigenfunctions $\phi_1,\phi_2,\ldots$ of the unweighted Laplace-Beltrami operator $\Delta = \Delta_{\nu}$.
+Then letting $\Phi_K = \mathrm{span}\{f_1,\ldots,f_K\}$, the least-squares estimator and test statistic
+\begin{equation*}
+\wt{f}_{\mathrm{LS}} := \argmin_{f \in \Phi_K} \|Y - f\|_n^2,\textrm{and}~~\wt{T}_{\mathrm{LS}} = \|\wt{f}\|_{\nu}^2
+\end{equation*}
+are still rate-optimal over $H_0^s(\mc{X})$.
+
 However, this is not a totally satisfactory fix.
-For one thing, the least squares approach still requires that we know $\Delta$, and is not well suited for the case where $\mc{X}$ is a manifold, and $\Delta$ is the manifold Laplace-Beltrami operator.
-Additionally, diagonalizing even the unweighted Laplace-Beltrami operator $\Delta$ is quite difficult for all but a few special domains, such as $\mc{X} = [0,1]^d$.
+For one thing, the least squares approach still requires that we know $\Delta$, and is not well suited for the case where $\mc{X}$ is an unknown manifold, and $\Delta$ is the manifold Laplace-Beltrami operator.
+Additionally, even if we know $\mc{X}$, diagonalizing the Laplace-Beltrami operator $\Delta$ is quite difficult for all but a few special domains, such as $\mc{X} = [0,1]^d$.
 
@@
 It simply reflects that in the standard nonparametric regression setup---which we adopt in the main part of this paper, and in which $P$ is assumed to be equivalent to Lebesgue measure---the biases of Laplacian eigenmaps and kernel smoothing are equivalent.
-There has been some work \textcolor{red}{(Rigollet, Wasserman, Niyogi, El Alaoui)} analyzing semi-supervised learning in the case where $f_0$ and $p$ satisfy some special relationship, such as the cluster assumption.
-However, a comprehensive analysis comparing the error of Laplacian based methods (e.g.
-eigenmaps) to that of other methods remains outstanding.
+On the other hand, when $f_0$ and $p$ satisfy some special relationship, such as the cluster assumption, the biases of these two methods can be quite different.
+There has been some work \textcolor{red}{(Rigollet, Wasserman, Niyogi, El Alaoui)} analyzing semi-supervised learning under various instantiations of the cluster assumption.
+However, a comprehensive analysis of various methods, including Laplacian eigenmaps, in this setting remains outstanding.
+
+\paragraph{Laplacian smoothing.}
+\textcolor{red}{(TODO)}
 
 \subsection{Future Work}
-As mentioned at the outset, our work can be viewed as a contribution both to the fields of nonparametric regression with series estimators, and to graph-based learning.
-Fittingly, we end our discussion by mentioning some open work in each of these directions.
+We view our work can be viewed as a contribution both to the fields of nonparametric regression with series estimators, and to graph-based learning.
+We end our discussion by mentioning some open work in each of these directions.
 
 Much is known about classical spectral series methods beyond their rate optimality.
-For instance, such estimators and tests exhibit \emph{sharp optimality}, meaning their risk can be bounded to within a $(1 + o(1))$ factor of the optimal risk.
-They can adapt to unknown smoothness of the regression function.
-They can be used to estimate smooth functionals of the regression function, or alternatively to form confidence sets in $L^2(P)$.
+An incomplete list: such estimators and tests exhibit \emph{sharp optimality}, meaning their risk is within a $(1 + o(1))$ factor of the optimal risk; they can adapt to unknown smoothness of the regression function; they can be used to estimate smooth functionals of the regression function;  finally, they can be used to form confidence sets in $L^2(P)$.
 It would be interesting to see if Laplacian eigenmaps could replicate the performance of classical methods in any, or all, of these problems.

### 2021-03-27 notes/weekly_notes/12.24.20.NotesForWeek.tex
Completed proof regarding bias of kernel smoothing.

@@
 \end{lemma}
-\begin{proof}[Proof (when dimension $d = 1$).]
-	\textcolor{red}{(TODO)}
-\end{proof}
+For the proof of this claim see Section~\ref{sec:kernel_smoothing_bias_pf}.
 
@@
 \section{Proof of Lemma~\ref{lem:kernel_smoothing_bias}}
+\label{sec:kernel_smoothing_bias_pf}
 Letting $\Omega_{\varepsilon} = \{x: B(x,\varepsilon) \subset \Omega\}$ and $\partial_{\varepsilon}\Omega = \Omega \setminus \Omega_{\varepsilon}$, we have that 
 \begin{equation*}
-\|L_{P,\varepsilon}u\|_P^2 = \underbrace{\int_{\Omega_{\varepsilon}} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_1} + \underbrace{\int_{\partial_{\varepsilon}\Omega} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_2}
+\|L_{P,\varepsilon}u\|_P^2 = \underbrace{\int_{\Omega_{\varepsilon}} \bigl(L_{P,\varepsilon}u(x)\bigr)^2 \,dP(x)}_{:=I_1} + \underbrace{\int_{\partial_{\varepsilon}\Omega} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_2}
 \end{equation*}
-We will separately upper bound $I_1$---using the higher order property of $\eta$---and $I_2$---using the zero-trace property of $u$ and its $1$st up through $j - 1$st order derivatives.
+We will separately upper bound $I_1$---using the higher order property of $\eta$---and $I_2$---using the zero-trace property of $u$ and its $1$st up through $(j - 1)$st order derivatives.
+
+\subsection{Upper bound on $I_1$.}
+\paragraph{Taylor expansions, and relevant facts.}
+Suppose $x \in \mc{X}_{\varepsilon}$ and $x' \in B(x,\varepsilon)$. We take a Taylor expansion of $u(x')$ around $x' = x$,
+\begin{equation*}
+u(x') - u(x) = \sum_{j = 1}^{s - 1} \frac{1}{j!}\bigl(d_x^{j}u\bigr)(x' - x) + r_{x'}^{s}(x;u)
+\end{equation*}
+We likewise take an order $k$ Taylor expansion of $p(x')$ around $x'  = x$ (where $k = 1$ if $s = 1$, and otherwise $k = s - 1$,)
+\begin{equation*}
+p(x') = p(x) + \sum_{j = 1}^{k - 1}\frac{1}{j!}\bigl(d_x^{j}p\bigr)(x' - x) + r_{x'}^{k}(x;p).
+\end{equation*}
+Here for a given $z \in \Rd$ and $j$-times differentiable function $f: \mc{X} \to \Reals$, we use the notation $\bigl(d_x^jf\bigr)(z) := \sum_{\abs{\alpha} = j} D^{\alpha}f(x) z^{\alpha}$.
+The remainder term $r_{x'}$ is given by
+\begin{equation*}
+r_{x'}^j(x;f) = \frac{1}{(j - 1)!} \int_{0}^{1}(1 - t)^{j - 1} \bigl(d_{x + t(x' - x)}^{j}f\bigr)(x' - x) \,dt,
+\end{equation*}
+where we note that $x + t(x' - x) \in B(x,\varepsilon) \subseteq \mc{X}$. Finally, we have adopted the convention that $\sum_{j = 1}^{0} a_j = 0$.
+
+We now recall some standard facts regarding Taylor expansions.
+First, remember that that $\bigl(d_x^{j}f\bigr)(z)$ is a degree-$j$ polynomial---and thus a $j$-homogeneous function---in $z$.
+On the other hand, for any degree-$j$ polynomial $q_j(z)$ and order-$s$ kernel $\eta$, if $j < s$ by converting to polar coordinates we can verify that
+\begin{equation*}
+\int_{B(0,1)} q_j(z) \eta(\|z\|) \,dz = 0.
+\end{equation*}
+We now give estimates on the remainder term in both sup-norm and $L^2(\mc{X}_{\varepsilon})$ norm, each of which hold for any $z \in B(0,1)$. In sup-norm, we have that 
+\begin{equation*}
+\sup_{x \in \mc{X}_{\varepsilon}}|r_{x + \varepsilon z}^j(x;f)| \leq C \varepsilon^j \|f\|_{C^j(\mc{X})},
+\end{equation*}
+whereas in $L^2(\mc{X}_{\varepsilon})$ norm we have,
+\begin{equation}
+\label{eqn:sobolev_remainder_term}
+\int_{\mc{X}_{\varepsilon}} \bigl|r_{x + t \varepsilon z}^j(x;f)\bigr|^2 \,dx \leq \varepsilon^{2j} \int_{\mc{X}_{\varepsilon}} \int_{0}^{1} |d_{x + t\varepsilon z}^jf(z)|^2 \,dt \,dx \leq \varepsilon^{2j} \|d^jf\|_{\Leb^2(\mc{X})}^2.
+\end{equation}
 
 \paragraph{Upper bound on $I_1$.}
+Taking a Taylor expansion of $u(x')$ around $x' = x$, we see that
+\begin{equation*}
+d_P(x) \cdot L_{P,\varepsilon}u(x) = \sum_{j = 1}^{s - 1} \int \bigl(d_x^ju\bigr)(x' - x) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') + \int r_{x'}^j(x;u) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x')
+\end{equation*}
+Next we take a Taylor expansion of $p(x')$ around $x' = x$ and represent $L_{P,\varepsilon}u(x)$ as the sum of three terms,
+\begin{equation}
+\label{pf:kernel_smoothing_bias_1}
+\begin{aligned}
+d_P(x) \cdot L_{P,\varepsilon}u(x) & = \sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{k - 1} \int \bigl(d_x^{j_1}u\bigr)(x' - x) \bigl(d_x^{j_2}p\bigr)(x' - x) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \quad + \\
+& \quad \sum_{j = 1}^{s - 1} \int \bigl(d_x^ju\bigr)(x' - x)  r_{x'}^{k}(x;p) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \quad  + \\
+& \quad \int r_{x'}^j(x;u) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \\
+& := J_1(x) + J_2(x) + J_3(x).
+\end{aligned}
+\end{equation}
+Note immediately that by Assumption~\ref{asmp:density}, we have that $\inf_{x \in \mc{X}_{\varepsilon}}|d_{P}(x)| \geq p_{\min} \varepsilon^d.$ We will additionally show that for each of $m = 1,2,3$, 
+\begin{equation*}
+\int_{\Omega_{\varepsilon}} |J_m(x)|^2 \,dP(x) \leq C \varepsilon^{2(s + d)} \|u\|_{H^s(\mc{X})}^2,
+\end{equation*}
+for a constant $C$ that does not depend on $u$. Plugging these bounds back into~\eqref{pf:kernel_smoothing_bias_1} will imply that
+\begin{equation*}
+I_1 \leq C \varepsilon^{2s} \|u\|_{H^s(\mc{X})}^2.
+\end{equation*}
+
+\underline{\textit{Upper bound on $J_1(x)$}}.
+Recalling that $d_x^jf(\cdot)$ is a $j$-homogeneous function, we have that
+\begin{align*}
+J_1(x) & = \sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{k - 1} \int_{\mc{X}} \bigl(d_x^{j_1}u\bigr)(x' - x) \cdot \bigl(d_x^{j_2}p\bigr)(x' - x) \cdot \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \\
+& = \varepsilon^d \sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{k - 1} \int_{B(0,1)} \bigl(d_{x }^{j_1}u\bigr)(\varepsilon z) \cdot \bigl(d_{x}^{j_2}p\bigr)(\varepsilon z) \cdot \eta\bigl(\|z\|\bigr) \,dz \\
+& = \varepsilon^d \sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{k - 1} \varepsilon^{j_1 + j_2} \int_{B(0,1)} \bigl(d_{x}^{j_1}u\bigr)(z) \cdot \bigl(d_{x}^{j_2}p\bigr)(z) \cdot \eta\bigl(\|z\|\bigr) \,dz.
+\end{align*}
+Noting $d_{x}^{j_1}u \cdot d_x^{j_2}p$ is a degree $j_1 + j_2$ polynomial, we note that for all terms in the above sum such that $j_1 + j_2 < s$,
+\begin{equation*}
+\int_{B(0,1)} \bigl(d_{x}^{j_1}u\bigr)(z) \cdot \bigl(d_{x}^{j_2}p\bigr)(z) \cdot \eta\bigl(\|z\|\bigr) \,dx' = 0.
+\end{equation*}
+On the other hand if $j_1 + j_2 \geq s$, we have that
+\begin{align*}
+\biggl|\int_{B(0,1)} \bigl(d_{x}^{j_1}u\bigr)(z) \cdot \bigl(d_{x}^{j_2}p\bigr)(z) \cdot \eta\bigl(\|z\|\bigr) \,dz\biggr| \leq \|d^jp\|_{L^{\infty}} \int_{B(0,1)} \Bigl|\bigl(d_x^{j}u\bigr)(z)\Bigr| \cdot \Bigl|\eta(\|z\|)\Bigr| \,dz.
+\end{align*}
+Consequently, applying these bounds for each $x \in \mc{X}_{\varepsilon}$ and integrating, we obtain
+\begin{align*}
+\int |J_1(x)|^2 \,dP(x) & \leq C \varepsilon^{2(s + d)} \|p\|_{C^k(\mc{X})} \sum_{j = 1}^{s - 1}  \int_{\mc{X}_{\varepsilon}} \biggl(\int_{B(0,1)} \Bigl|\bigl(d_x^{j}u\bigr)(z)\Bigr| \cdot \Bigl|\eta(\|z\|)\Bigr| \,dz\biggr)^2 \,dx \\
+& \leq C \varepsilon^{2(s + d)} \|p\|_{C^k(\mc{X})} \sum_{j = 1}^{s - 1} \|d_ju\|_{L^2(\mc{X}_{\varepsilon})}^2 \\
+& \leq  C \varepsilon^{2(s + d)} \|p\|_{C^k(\mc{X})} \|u\|_{H^s(\mc{X})}^2,
+\end{align*}
+where we show the middle inequality in~\eqref{pf:kernel_smoothing_bias_3}, below.
+
+\underline{\textit{Upper bound on $J_2(x)$}}.
+Recalling that $|r_{x + z\varepsilon}^{k}(x; p)| \leq C\varepsilon^{k}\|p\|_{C^{k}(\mc{X})}$ for any $z \in B(0,1)$, and that $d_x^jf(\cdot)$ is a $j$-homogeneous function, we have that
+\begin{align}
+|J_2(x)| & \leq \sum_{j = 1}^{s - 1} \frac{1}{j!}\int \Bigl|\bigl(d_x^{j}u\bigr)(x' - x)\Bigr| \cdot \Bigl|r_{x'}^{k}(x;p)\Bigr| \cdot \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr)\biggr| \,dx' \nonumber \\
+& = \varepsilon^{d} \sum_{j = 1}^{s - 1} \frac{1}{j!}\int_{B(0,1)} \Bigl|\bigl(d_x^{j}u\bigr)(\varepsilon z)\Bigr| \cdot |r_{x + z\varepsilon}^{k}(x;p)| \cdot \Bigl|\eta(\|z\|)\Bigr| \,dz \nonumber \\
+& \leq C\varepsilon^{k + 1 + d}\|p\|_{C^{k}(\mc{X})} \sum_{j = 1}^{s - 1} \frac{1}{j!} \int_{B(0,1)} \Bigl|\bigl(d_x^{j}u\bigr)(z)\Bigr| \cdot \Bigl|\eta(\|z\|)\Bigr| \,dz \label{pf:kernel_smoothing_bias_2}
+\end{align}
+Furthermore, for each $j = 1,\ldots,s - 1$ convolution of $d_x^ju$ with $\eta$ increases the $\Leb^2(\mc{X}_{\varepsilon})$ norm by at most a constant, meaning 
+\begin{equation}
+\label{pf:kernel_smoothing_bias_3}
+\begin{aligned}
+\int_{\mc{X}_{\varepsilon}} \biggl(\int_{B(0,1)} \Bigl|\bigl(d_x^{j}u\bigr)(z)\Bigr| \cdot \Bigl|\eta(\|z\|)\Bigr| \,dz\biggr)^2 \,dx & \leq \int_{\mc{X}_{\varepsilon}} \biggl(\int_{B(0,1)} \Bigl[\bigl(d_x^ju\bigr)(z)\Bigr]^2 \,dz \biggr) \cdot \biggl(\int_{B(0,1)} K(\|z\|) \,dz \biggr) \,dx \\
+& \leq \|\eta\|_{\infty}^2  \int_{B(0,1)} \int_{\mc{X}_{\varepsilon}} \Bigl[\bigl(d_x^ju\bigr)(z)\Bigr]^2 \,dx \,dz \\
+& \leq \|\eta\|_{\infty}^2 \nu_d \cdot \|d^ju\|_{\Leb^2(\mc{X_{\varepsilon}})}^2.
+\end{aligned}
+\end{equation}
+Combining this with~\eqref{pf:kernel_smoothing_bias_2}, we conclude that
+\begin{align*}
+\int_{\mc{X}_{\varepsilon}} |J_2(x)|^2 \,dP(x) & \leq C p_{\max} \|p\|_{C^{k}(\mc{X})}^2 \varepsilon^{2(k + 1 + d)} \sum_{j = 1}^{s - 1} \biggl(\frac{1}{j!} \int_{B(0,1)} \Bigl|\bigl(d_x^{j}u\bigr)(z)\Bigr| \cdot \Bigl|\eta(\|z\|)\Bigr| \,dz\biggr) \\
+& \leq  C p_{\max} \|p\|_{C^{k}(\mc{X})}^2 \|\eta\|_{\infty}^2 \nu_d \varepsilon^{2(k + 1 + d)} \sum_{j = 1}^{s - 1} \cdot \|d^ju\|_{\Leb^2(\mc{X_{\varepsilon}})}^2 \\
+& \leq C \varepsilon^{2(k + 1 + d)} \|u\|_{H^s(\mc{X})}^2,
+\end{align*}
+whence the claim follows since by assumption either $k + 1 > s$ (when $s = 1$), or $k + 1 = s$ (when $s \geq 2$).
+
+\underline{\textit{Upper bound on $J_3(x)$}}.
+We begin with a pointwise upper bound on $|J_3(x)|$, 
+\begin{align*}
+|J_3(x)| & \leq p_{\max} \int_{\mc{X}} \bigl|r_{x'}^s(x;u)\bigr| \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr)\biggr| \,dx' \\
+& = p_{\max} \varepsilon^d \int_{B(0,1)} \bigl|r_{x + \varepsilon z}^s(x;u)\bigr| \bigl|\eta(\|z\|)\bigr| \,dz,
+\end{align*}
+and applying the Cauchy-Schwarz inequality, we deduce a pointwise upper bound on $|J_3(x)|^2$,
+\begin{align*}
+|J_3(x)|^2 & \leq p_{\max}^2 \varepsilon^{2d} \biggl(\int_{B(0,1)} \bigl|r_{x + \varepsilon z}^s(x;u)\bigr|^2 \,dz\biggr) \cdot \biggl(\int_{B(0,1)} \bigl|\eta(\|z\|)\bigr| \,dz\biggr) \\
+& \leq p_{\max}^2 \|\eta\|_{\infty}^2 \nu_d \varepsilon^{2d}  \int_{B(0,1)} \bigl|r_{x + \varepsilon z}^s(x;u)\bigr|^2 \,dz.
+\end{align*}
+Applying this pointwise over all $x \in \mc{X}_{\varepsilon}$ and integrating, we obtain
+\begin{align*}
+\int_{\mc{X}_{\varepsilon}} |J_3(x)|^2 \,dP(x) & \leq p_{\max}^3 \|\eta\|_{\infty}^2 \nu_d \varepsilon^{2d} \int_{\mc{X}_{\varepsilon}} \int_{B(0,1)} \bigl|r_{x + \varepsilon z}^s(x;u)\bigr|^2 \,dz \\
+& = p_{\max}^3 \|\eta\|_{\infty}^2 \nu_d \varepsilon^{2d} \int_{B(0,1)} \int_{\mc{X}_{\varepsilon}} \bigl|r_{x + \varepsilon z}^s(x;u)\bigr|^2 \,dz \\
+& \leq p_{\max}^3 \|\eta\|_{\infty}^2 \nu_d^2 \varepsilon^{2(s + d)} \|d^sf\|_{L^2(\mc{X}_{\varepsilon})},
+\end{align*}
+with the last inequality following from~\eqref{eqn:sobolev_remainder_term}.
 
 \paragraph{Upper bound on $I_2$.}
+We begin by demonstrating that $I_2$ is upper bounded by $C\|u\|_{\Leb^2(\partial_{2\varepsilon}\mc{X})}^2$.
+Recall that by assumption $d_P(x) \geq \alpha p_{\min} \varepsilon^d$ for each $x \in \mc{X}$.
+On the other hand, at every $x \in \mc{X}$ we have that
+\begin{align*}
+\bigl|d_P(x) \cdot L_{P,\varepsilon}u(x)\bigr|^2 & \leq p_{\max}^2 \biggl[\int \Bigl(|u(x)|^2 + |u(x')|^2\Bigr) K^{1/2}\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx'\biggr] \cdot \biggl[\int K^{1/2}\biggl(\frac{\|x' - x|}{\varepsilon}\biggr) \,dx\biggr] \\
+& \leq 2 p_{\max}^2 \|\eta\|_{\infty} \varepsilon^d \int \Bigl(|u(x)|^2 + |u(x')|^2\Bigr) K^{1/2}\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx'.
+\end{align*}
+It follows that
+\begin{align*}
+I_2 & = \int_{\partial_{\varepsilon}\mc{X}} \bigl|L_{P,\varepsilon}u(x)\bigr|^2 \,dP(x) \\
+& \leq \frac{2p_{\max}^2 |\eta\|_{\infty}}{p_{\min}^2 \alpha \varepsilon^d} \int_{\partial_{\varepsilon}\mc{X}} \int_{\mc{X}} \Bigl(|u(x)|^2 + |u(x')|^2\Bigr) K^{1/2}\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \\
+& \leq \frac{4p_{\max}^2 \|\eta\|_{\infty}^2}{p_{\min}^2 \alpha} \int_{\partial_{2\varepsilon}\mc{X}} |u(x)|^2 \,dx.
+\end{align*}
+We will now use the zero-trace property of $u$ to show that there exists a number $c$, which depends only on $\mc{X}$ such that for all $\varepsilon \leq c$,
+\begin{equation}
+\label{pf:kernel_smoothing_bias_5}
+\int_{\partial_{\varepsilon}\mc{X}} |u(x)|^2 \,dx \leq C\varepsilon^{2s}.
+\end{equation}
+We will build to this result by a series of intermediate steps. Many of these steps are related to the proof of Theorem 18.1 in \textcolor{red}{(Leoni)}.
+
+\underline{\textit{Step 1: Local Patch.}}
+To begin, we assume that for some $c_0 > 0$ and a Lipschitz mapping $\phi: \Reals^{d - 1} \to [-c_0,c_0]$, we have that $u \in C_c^{\infty}(U_{\phi}(c_0))$, where 
+\begin{equation*}
+U_{\phi}(c_0) = \Bigl\{y \in Q(0,c_0): \phi(y_{-d}) \leq y_d\Bigr\}, 
+\end{equation*}
+and here $Q(0,c_0)$ is the $d$-dimensional cube of side length $c_0$, centered at $0$.
+We will show that for all $0 < \varepsilon < c_0$, and for the tubular neighborhood $V_{\phi}(\varepsilon) = \{y \in Q(0,c_0): \phi(y_{-d}) \leq y_d \leq \phi(y_{-d}) + \varepsilon\}$, we have that
+\begin{equation*}
+\int_{V_{\phi}(\varepsilon)} |u(x)|^2 \,dx \leq C\varepsilon^{2s} \|u\|_{H^s(U_{\phi}(c_0))}^2.
+\end{equation*}
+For a given $y = (y',y_d) \in V_{\phi}(\varepsilon)$, let $y_0 = (y',\phi(y'))$.
+Taking the Taylor expansion of $u(y)$ around $y = y_0$ because $u$ is compactly supported in $V_{\phi}$ it follows that,
+\begin{align*}
+u(y) & = u(y_0) + \sum_{j = 1}^{s - 1} \frac{1}{j!} D^{je_d}u(y_0) \bigl(y_d - \phi(y')\bigr)^j + \frac{1}{(s - 1)!}\int_{\phi(y')}^{y_d} (1 - t)^{s - 1} D^{se_d}u(y',z) \bigl(y_d - z\bigr)^{s - 1} \,dz \Longrightarrow\\
+|u(y)| & \leq C\varepsilon^{s - 1}\int_{\phi(y')}^{y_d} \bigl|D^{se_d}u(y',z)\bigr| \,dz. 
+\end{align*}
+Consequently, by squaring both sides and applying Cauchy-Schwarz, we have that
+\begin{equation*}
+|u(y)|^2 \leq C\varepsilon^{2(s - 1)} \biggl(\int_{\phi(y')}^{y_d} \bigl|D^{se_d}u(y',z)\bigr| \,dz\biggr)^2 \leq C\varepsilon^{2s - 1} \int_{\phi(y')}^{y_d} \bigl|D^{se_d}u(y',z)\bigr|^2 \,dz.
+\end{equation*}
+Applying this bound for each $y \in V_{\phi}(\varepsilon)$, and then integrating, we obtain
+\begin{align}
+\int_{V_{\phi}(\varepsilon)} |u(y)|^2 \,dy & \leq \int_{Q_{d - 1}(c_0)} \int_{\phi(y')}^{\phi(y') + \varepsilon} |u(y',y_d)|^2 \,dy_d \,dy' \nonumber \\
+& \leq C\varepsilon^{2s - 1}\int_{Q_{d - 1}(c_0)}  \int_{\phi(y')}^{\phi(y') + \varepsilon} \int_{\phi(y')}^{y_d} \bigl|D^{se_d}u(y',z)\bigr|^2 \,dz \,dy_d \,dy' \label{pf:kernel_smoothing_bias_4}
+\end{align}
+where we have written $Q_{d - 1}(0,c_0)$ for the $d - 1$ dimensional cube of side length $c_0$, centered at $0$. Exchanging the order of the inner two integrals then gives
+\begin{align*}
+\int_{\phi(y')}^{\phi(y') + \varepsilon} \int_{\phi(y')}^{y_d} \bigl|D^{se_d}u(y',z)\bigr|^2 \,dz \,dy_d & = \int_{\phi(y')}^{\phi(y') + \varepsilon} \int_{z}^{\varepsilon} \bigl|D^{se_d}u(y',z)\bigr|^2 \,dy_d \,dz \\
+& \leq C \varepsilon \int_{\phi(y')}^{\phi(y') + \varepsilon} \bigl|D^{se_d}u(y',z)\bigr|^2 \,dz \\
+& \leq C \varepsilon \int_{\phi(y')}^{c_0} \bigl|D^{se_d}u(y',z)\bigr|^2 \,dz.
+\end{align*}
+Finally, plugging back into~\eqref{pf:kernel_smoothing_bias_4}, we conclude that
+\begin{equation*}
+\int_{V_{\phi}(\varepsilon)} |u(y)|^2 \,dy \leq C \varepsilon^{2s} \int_{Q_{d - 1}(0,c_0)} \int_{\phi(y')}^{c_0} \bigl|D^{se_d}u(y',z)\bigr|^2 \,dz \,dy' \leq C \varepsilon^{2s} |u|_{H^s(U_{\phi}(c_0))}^2.
+\end{equation*}
+
+\underline{\textit{Step 2: Rigid motion of local patch.}} Now, suppose that at a point $x_0 \in \partial \mc{X}$, there exists a rigid motion $T: \Rd \to \Rd$ for which $T(x_0) = 0$, and a number $C_0$ such that for all $\varepsilon \cdot C_0 \leq c_0$,
+\begin{equation*}
+T\bigl(Q_{T}(x_0,c_0) \cap \partial_{\varepsilon}\mc{X}\bigr) \subseteq V_{\phi}\bigl(C_0\varepsilon\bigr) \quad\textrm{and}\quad T\bigl(Q_T(x_0,c_0) \cap \mc{X}\bigr) = U_{\phi}(c_0).
+\end{equation*}
+Here $Q_{T}(x_0,c_0))$ is a (not necessarily coordinate-axis-aligned) cube of side length $c_0)$, centered at $x_0$.
+Define $v(y) := u(T^{-1}(y))$ for $y \in U_{\phi}(c_0)$.
+If $u \in C_c^{\infty}(\mc{X})$, then $v \in C_c^{\infty}(U_{\phi}(c_0))$, and moreover $\|v\|_{H^s(U_{\phi}(c_0))}^2 = \|u\|_{H^s(Q_{T}(x_0,c_0) \cap \mc{X})}^2$.
+Therefore, using the upper bound that we derived in Step 1,
+\begin{equation*}
+\int_{V_{\phi}(C_0 \cdot \varepsilon)} |v(y)|^2 \,dy \leq C \varepsilon^{2s} \|v\|_{H^s(U_{\phi}(c_0))}^2,
+\end{equation*}
+we conclude that
+\begin{align*}
+\int_{Q_{T}(x_0,c_0) \cap \partial_{\varepsilon}\mc{X}} |u(x)|^2 \,dx & = \int_{T(Q_T(x_0,c_0)) \cap \partial_{\varepsilon}\mc{X})} |v(y)|^2 \,dy \\
+& \leq \int_{V_{\phi}(C_0 \cdot \varepsilon)} |v(y)|^2 \,dy \\
+& \leq C \varepsilon^{2s} \|v\|_{H^s(U_{\phi}(c_0))}^2 = C \varepsilon^{2s} \|u\|_{H^s(Q_{T}(x_0,c_0)) \cap \mc{X})}^2 \leq C \varepsilon^{2s} \|u\|_{H^s(\mc{X})}^2.
+\end{align*}
+
+\underline{\textit{Step 3: Lipschitz domain}}.
+Finally, we deal with the case where $\mc{X}$ is assumed to be an open, bounded subset of $\Rd$, with Lipschitz boundary.
+In this case, at every $x_0 \in \partial \mc{X}$, there exists a rigid motion $T_{x_0}: \Rd \to \Rd$ such that $T_{x_0}(x_0) = 0$, a number $c_0(x_0)$, a Lipschitz function $\phi_{x_0}:\Reals^{d - 1} \to [-c_0,c_0]$, and a number $C_0(x_0)$, such that for all $\varepsilon \cdot C_0(x_0) \leq c_0(x_0)$,
+\begin{equation*}
+T\bigl(Q_{T}(x_0,c_0(x_0)) \cap \partial_{\varepsilon}\mc{X}\bigr) \subseteq V_{\phi}\bigl(C_0(x_0) \cdot \varepsilon\bigr) \quad\textrm{and}\quad T\bigl(Q_T(x_0,c_0(x_0)) \cap \mc{X}\bigr) = U_{\phi}(c_0(x_0)).
+\end{equation*}
+Therefore for every $x_0 \in \partial \mc{X}$, it follows from the previous step that
+\begin{equation*}
+\int_{Q_{T_{x_0}}(x_0,c_0(x_0)) \cap \partial_{\varepsilon}\mc{X}} |u(x)|^2 \,dx \leq C(x_0) \varepsilon^{2s} \|u\|_{H^s(\mc{X})}^2,
+\end{equation*}
+where on the right hand side $C(x_0)$ is a constant that may depend on $x_0$, but not on $u$ or $\varepsilon$.
+
+We conclude by taking a collection of cubes that covers $\partial_{\varepsilon}\mc{X}$ for all $\epsilon$ sufficiently small.
+First, we note that by a compactness argument there exists a finite subset of the collection of cubes $\{Q_{T_{x_0}}(x_0,c_0(x_0)/2): x_0 \in \partial\mc{X} \}$ which covers $\partial \mc{X}$, say $Q_{T_{x_1}}(x_1,c_0(x_1)/2),\ldots, Q_{T_{x_N}}(x_N,c_0(x_N)/2)$.
+Then, for any $\varepsilon \leq \min_{i = 1,\ldots,N} c_0(x_i)/2$, it follows from the triangle inequality that
+\begin{equation*}
+\partial_{\varepsilon}\mc{X} \subseteq \bigcup_{i = 1}^{N} Q_{T_{x_i}}(x_i, c_0(x_i)).
+\end{equation*}
+As a result,
+\begin{equation*}
+\int_{\partial_{\varepsilon}\mc{X}} |u(x)|^2 \leq \sum_{i = 1}^{N} \int_{Q_{T_{x_i}}(x_i, c_0(x_i)) \cap \partial_{\varepsilon}(\mc{X})} |u(x)|^2 \leq  \varepsilon^{2s} \|u\|_{H^s(\mc{X})}^2 \sum_{i = 1}^{N}C_0(x_i),
+\end{equation*}
+which proves the claim of~\eqref{pf:kernel_smoothing_bias_5}.
 

### 2021-03-29 manuscripts/laplacian_eigenmaps/appendix.tex
Finished editting higher-order section of Laplacian eigenmaps paper.
Added organization for laplacian eigenmaps proofs.

@@
 		\begin{equation*}
-		\sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 \geq 16\frac{\sqrt{2K}}{n}\biggl(\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr).
+		\sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 \geq 16\frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr].
 		\end{equation*}
@@
 
-\section{Proofs in Manifold Setting}
-\label{sec:manifold_proofs}
+\section{Graph quadratic form, flat Euclidean domain}
+\label{sec:graph_quadratic_form_euclidean}
+
+\section{Graph quadratic form, manifold domain}
+\label{sec:graph_quadratic_form_manifold}
 
@@
 
-\section{Comparison of Laplacian eigenmaps and kernel smoothing in non-uniform density setting}
+\section{Lower bound on empirical norm}
+\label{sec:empirical_norm}
+In this Section we prove Proposition~\ref{prop:empirical_norm_sobolev} (in Section~\ref{subsec:empirical_norm_sobolev}).
+We also prove an analogous result when $\mc{X}$ is a manifold as in Model~\ref{def:model_manifold} (in Section~\ref{subsec:empirical_norm_sobolev_manifold}), and we prove some sharper estimates in the case where the function is bounded in H\"{o}lder norm, in Section~\ref{subsec:empirical_norm_holder}.
+
+\subsection{Proof of Proposition~\ref{prop:empirical_norm_sobolev}}
+\label{subsec:empirical_norm_sobolev}
+\textcolor{red}{(TODO)}
+
+\subsection{Analysis of empirical norm in manifold setting}
+\label{subsec:empirical_norm_soboolev_manifold}
+
+\subsection{Better bounds assuming H\"{o}lder smoothness}
+\label{subsec:empirical_norm_holder}
+
+\section{Analysis of kernel smoothing}
+\label{subsec:kernel_smoothing}
+In this section we prove Lemma~\ref{lem:kernel_smoothing_laplacian_eigenmaps} (in Section~\ref{subsec:pf_kernel_smoothing_laplacian_eigenmaps}) and Lemma~\ref{lem:kernel_smoothing_bias} (in Section~\ref{subsec:pf_kernel_smoothing_bias}).
+In Section~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we give a sequence $\{p^{(n)}(x), f_0^{(n)}(x)\}_{n \in \mathbb{N}}$ for which the estimator $T_{h,n}\wc{f}$ (kernel extension of Laplacian eigenmaps) dominates direct kernel smoothing of the responses, in the sense that
+\begin{equation*}
+\lim_{n \to \infty} \sup_{h'} \frac{\|T_{h_n,n}\wc{f} - f_0\|_P^2}{\|T_{h',n}Y - f_0\|_P^2 } = 0,\quad \textrm{with probability $1$.}
+\end{equation*}
+
+\subsection{Proof of Lemma~\ref{lem:kernel_smoothing_laplacian_eigenmaps}}
+\label{subsec:pf_kernel_smoothing_laplacian_eigenmaps}
+
+\subsection{Proof of Lemma~\ref{lem:kernel_smoothing_bias}}
+\label{subsec:pf_kernel_smoothing_bias}
+
+\subsection{Comparison of Laplacian eigenmaps to kernel smoothing}
+\label{subsec:eigenmaps_beats_kernel_smoothing}
 
@@
 \end{itemize}
+
+\section{Miscellaneous}
+Here we give some helpful Lemmas used at various points in the above proofs. We also prove that spectral sparsification does not change the rate of convergence of $\wh{f}$.

### 2021-03-31 manuscripts/laplacian_eigenmaps/appendix.tex
Added proof of pointwise convergence of iterated non-local Laplacian.

@@
 
-\section{Graph quadratic form, flat Euclidean domain}
+\section{Graph Sobolev semi-norm, flat Euclidean domain}
 \label{sec:graph_quadratic_form_euclidean}
+In this section we prove Proposition~\ref{prop:graph_seminorm_ho}. The proposition will follow from several intermediate results.
+\begin{enumerate}
+	\item~In Section~\ref{subsec:decomposition_graph_seminorm}, we show that
+	\begin{equation}
+	\label{pf:graph_seminorm_ho_1}
+	\dotp{L_{n,\varepsilon}^sf}{f}_n \leq \frac{1}{\delta} \dotp{L_{P,\varepsilon}^sf}{f}_{P} + \frac{C\varepsilon^2}{n\varepsilon^{2 + d}}M^2.
+	\end{equation}
+	with probability at least $1 - \delta$.
+
+	We term the first term on the right hand side the \emph{non-local Sobolev semi-norm}, as it is a kernelized approximation to the Sobolev semi-norm $\dotp{\Delta_P^sf}{f}_{P}$.
+The second term on the right hand side is a pure bias term, which as we will see is negligible compared to the non-local quadratic form as long as $\varepsilon \ll n^{-1/(2(s -1 + d))}$.
+	\item~In Section~\ref{subsec:approximation_error_nonlocal_laplacian}, we show that when $x$ is sufficiently in the interior of $\mc{X}$, then $L_{P,\varepsilon}^kf(x)$ is a good approximation to $\Delta_P^kf(x)$, as long as $f \in H^{s}(\mc{X})$ and $p \in C^{s - 1}(\mc{X})$ for some $s \geq 2k + 1$.
+	\item~In Section~\ref{subsec:boundary_behavior_nonlocal_laplacian}, we show that when $x$ is sufficiently near the boundary of $\mc{X}$, then $L_{P,\varepsilon}^kf(x)$ is close to $0$, as long as $f \in H_0^{s}(\mc{X})$ for some $s \geq 2k$.
+	\item~In Section~\ref{subsec:estimate_nonlocal_seminorm}, we use the results of the preceding two sections to show that if $f \in H_0^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$, there exists a constant $C$ which does not depend on $f$ such that
+	\begin{equation}
+	\label{pf:graph_seminorm_ho_2}
+	\dotp{L_{P,\varepsilon}^sf}{f}_{P} \leq CM^2.
+	\end{equation}
+\end{enumerate}
+Finally, in Section~\ref{subsec:integrals} we provide some assorted estimates used in Sections~\ref{subsec:decomposition_graph_seminorm}.
+
+\paragraph{Proof (of Proposition~\ref{prop:graph_seminorm_ho}).}
+Proposition~\ref{prop:graph_seminorm_ho} follows immediately from~\eqref{pf:graph_seminorm_ho_1} and~\eqref{pf:graph_seminorm_ho_2}.
+
+\subsection{Decomposition of graph Sobolev semi-norm}
+\label{subsec:decomposition_graph_seminorm}
+\textcolor{red}{(TODO)}
+
+\subsection{Approximation error of non-local Laplacian}
+\label{subsec:approximation_error_nonlocal_laplacian}
+
+In this section, we establish the convergence $L_{P,\varepsilon}^kf \to \sigma_{\eta}^k\Delta_P^kf$ as $\varepsilon \to 0$.
+More precisely, we give an upper bound on the squared difference between $L_{P,\varepsilon}^kf$ and  $\sigma_{\eta}^k\Delta_P^kf$ as a function of $\varepsilon$.
+The bound holds for all $x \in \mc{X}_{k\varepsilon}$, and $f \in H^{s}(\mc{X})$, as long as $s \geq 2k + 1$.
+\textcolor{red}{(TODO): Make sure that $\sigma_{\eta}$ is defined somewhere.}
+\begin{lemma}
+	\label{lem:approximation_error_nonlocal_laplacian}
+	Assume Model~\ref{def:model_flat_euclidean}.
+Let $s \in \mathbb{N} \setminus \{0,1\}$, suppose that $f \in H^s(\mc{X};M)$, and if $s > 1$ suppose that $p \in C^{s - 1}(\mc{X})$.
+Let $L_{P,\varepsilon}$ be define with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_flat_euclidean}.
+Then there exist constants $C_1$ and $C_2$ that do not depend on $f$, such that each of the following statements hold.
+	\begin{itemize}
+		\item If $s$ is odd and $k = (s - 1)/2$, then
+		\begin{equation}
+		\label{eqn:approximation_error_nonlocal_laplacian_1}
+		\|L_{P,\varepsilon}^kf - \Delta_P^kf\|_{L^2(\mc{X}_{k\varepsilon})} \leq C_1 M \varepsilon
+		\end{equation}
+		\item If $s$ is even and $k = (s - 2)/2$, then
+		\begin{equation}
+		\label{eqn:approximation_error_nonlocal_laplacian_2}
+		\|L_{P,\varepsilon}^kf - \Delta_P^kf\|_{L^2(\mc{X}_{k\varepsilon})} \leq C_2 M \varepsilon^2.
+		\end{equation}
+	\end{itemize}
+\end{lemma}
+We remark that when $k = 1$ and $f \in C^3(\mc{X})$ or $C^4(\mc{X})$, statements of this kind are well known \textcolor{red}{(references)}, and indeed stronger results---with $L^{\infty}(\mc{X})$ norm replacing $L^2(\mc{X})$ norm---hold.
+When dealing with the iterated Laplacian, and functions $f$ which are regular only in the Sobolev sense, the proof is somewhat more lengthy, but the spirit of the result is largely the same.
+
+\paragraph{Proof (of Lemma~\ref{lem:approximation_error_nonlocal_laplacian}).}
+Throughout this proof, we shall assume that $f$ and $p$ are smooth functions, meaning they belong to $C^{\infty}(\mc{X})$.
+This is without loss of generality, since $C^{\infty}(\mc{X})$ is dense in both $H^s(\mc{X})$ and $C^{s - 1}(\mc{X})$, and since both sides of the inequalities~\eqref{eqn:approximation_error_nonlocal_laplacian_1} and~\eqref{eqn:approximation_error_nonlocal_laplacian_2} are continuous with respect to $\|\cdot\|_{H^s(\mc{X})}$ and $\|\cdot\|_{C^{s - 1}(\mc{X})}$ norms.
+
+We will actually prove a more general set of statements than contained in Lemma~\ref{lem:approximation_error_nonlocal_laplacian}, more general in the sense that they give estimates for all $k$, rather than simply the particular choices of $k$ given above.
+In particular, we will prove that the following two statements hold for any $s \in \mathbb{N}$ and any $k \in \mathbb{N} \setminus \{0\}$.
+\begin{itemize}
+	\item If $k \geq s/2$, then for every $x \in \mc{X}_{k\varepsilon}$, 
+	\begin{equation}
+	\label{pf:approximation_error_nonlocal_laplacian_0}
+	L_{P,\varepsilon}^kf(x) = g_s(x) \varepsilon^{s - 2k}
+	\end{equation}
+	for a function $g_s$ that satisfies
+	\begin{equation}
+	\label{pf:approximation_error_nonlocal_laplacian_0.5}
+	\|g_s\|_{L^2(\mc{X}_{k\varepsilon})} \leq C \|p\|_{C^{q}(\mc{X})}^k M 
+	\end{equation}
+	where $q = 1$ if $s =0$ or $s = 1$, and otherwise $q = s - 1$. 
+	\item If $k < s/2$, then for every $x \in \mc{X}_{k\varepsilon}$,
+	\begin{equation}
+	\label{pf:approximation_error_nonlocal_laplacian_1}
+	L_{P,\varepsilon}^kf(x) = \sigma_{\eta}^k \cdot \Delta_{P}^kf(x) + \sum_{j = 1}^{\floor{(s - 1)/2} - k} g_{2(j + k)}(x)\varepsilon^{2j} + g_{s}(x) \varepsilon^{s - 2k}.
+	\end{equation}
+	for functions $g_j$ that satisfy
+	\begin{equation}
+	\label{pf:approximation_error_nonlocal_laplacian_1.5}
+	\|g_j\|_{H^{s - j}(\mc{X}_{k\varepsilon})} \leq C \|p\|_{C^{s - 1}(\mc{X})}^k M.
+	\end{equation}
+\end{itemize}
+In the statement above, recall that $H^0(\mc{X}_{k\varepsilon}) = L^2(\mc{X}_{k\varepsilon})$.
+Additionally, note that we may speak of the pointwise behavior of derivatives of $f$ because we have assumed that $f$ is a smooth function.
+Observe that~\eqref{eqn:approximation_error_nonlocal_laplacian_1} follows upon taking $k = \floor{(s - 1)/2}$ in~\eqref{pf:approximation_error_nonlocal_laplacian_1}, whence we have
+\begin{equation*}
+\bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_{P}^kf(x)\bigr)^2 = \varepsilon^2 \bigl(g_s(x)\bigr)^2
+\end{equation*}
+for some $g_s \in \Leb^2(\mc{X}_{k\varepsilon},C \cdot M \cdot \|p\|_{C^{s - 1}(\mc{X})})$, and integrating over $\mc{X}_{k\varepsilon}$ gives the desired result.
+\eqref{eqn:approximation_error_nonlocal_laplacian_2} follows from~\eqref{pf:approximation_error_nonlocal_laplacian_1} in an identical fashion.
+
+It thus remains establish~\eqref{pf:approximation_error_nonlocal_laplacian_1}, and~\eqref{pf:approximation_error_nonlocal_laplacian_0} which is an important part of proving~\eqref{pf:approximation_error_nonlocal_laplacian_1}.
+We will do so by induction on $k$.
+Note that throughout, we will let $g_j$ refer to functions which may change from line to line, but which always satisfy~\eqref{pf:approximation_error_nonlocal_laplacian_1.5}.
+
+\underline{\textit{Proof of~\eqref{pf:approximation_error_nonlocal_laplacian_0} and~\eqref{pf:approximation_error_nonlocal_laplacian_1}, base case.}}
+
+We begin with the base case, where $k = 1$.
+Again, we point out that although desired result is known when $s = 3$ or $s = 4$, and $f$ is regular in the H\"{o}lder sense, we require estimates for all $s \in \mathbb{N}$ when $f$ is regular in the Sobolev sense.
+
+When $s = 0$, the inequality~\eqref{pf:approximation_error_nonlocal_laplacian_0} is implied by Lemma~\ref{lem:l2estimate_nonlocal_laplacian}.
+When $s \geq 1$, we proceed using Taylor expansion.
+For any $x \in \mc{X}_{\varepsilon}$, we have that $B(x,\varepsilon) \subseteq \mc{X}$.
+Thus for any $x' \in B(x,\varepsilon)$, we may take an order $s$ Taylor expansion of $f$ around $x' = x$, and an order $q$ Taylor expansion of $p$ around $x' = x$, where $q = 1$ if $s = 1$, and otherwise $q = s - 1$.
+(See Section~\ref{subsec:taylor_expansion} for a review of the notation we use for Taylor expansions, as well as some properties that we make use of shortly.) This allows us to express $L_{P,\varepsilon}f(x)$ as the sum of three terms,
+\begin{align*}
+L_{P,\varepsilon}f(x) & = \frac{1}{\varepsilon^{d + 2}}\sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1}\frac{1}{j_1!j_2!}  \int_{\mc{X}} \bigl(d_x^{j_1}f\bigr)(x' - x) \bigl(d_x^{j_2}p\bigr)(x' - x) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \quad + \\
+& \quad \frac{1}{\varepsilon^{d + 2}}\sum_{j = 1}^{s - 1} \frac{1}{j!} \int_{\mc{X}} \bigl(d_x^jf\bigr)(x' - x)  r_{x'}^{q}(x;p) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \quad  + \\
+& \quad \frac{1}{\varepsilon^{d + 2}} \int_{\mc{X}} r_{x'}^j(x;f) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \\
+& := G_1(x) + G_2(x) + G_3(x).
+\end{align*}
+We now separately consider each of $G_1(x),G_2(x)$ and $G_3(x)$. We will establish that if $s = 1$ or $s = 2$, then $G_1(x) = 0$, and otherwise if $s \geq 3$ that
+\begin{equation*}
+G_1(x) = \sigma_{\eta}\Delta_Pf(x) + \sum_{j = 1}^{\floor{(s - 1)/2} - 1}g_{2(j + 1)}(x)\varepsilon^{2j} + g_{s}(x)\varepsilon^{s - 2}.
+\end{equation*}
+On the other hand, we will establish that if $s = 1$ then $G_2(x) = 0$, and otherwise for $s \geq 2$
+\begin{equation*}
+\|G_2\|_{\Leb^2(\mc{X}_{\varepsilon})} \leq C \varepsilon^{s - 2} M \|p\|_{C^{s - 1}(\mc{X})};
+\end{equation*}
+this same estimate will hold for $G_3$ for all $s \geq 1$. Together these will imply~\eqref{pf:approximation_error_nonlocal_laplacian_0} and~\eqref{pf:approximation_error_nonlocal_laplacian_1}.
+
+\emph{Estimate on $G_1(x)$.}
+If $s = 1$, then $s - 1 = 0$, and so $G_1(x) = 0$. We may therefore suppose $s \geq 2$. In this case, a change of variables gives
+\begin{align}
+L_{P,\varepsilon}f(x) & = \frac{1}{\varepsilon^{d + 2}} \sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1} \frac{1}{j_1!j_2!}  \int_{\mc{X}} \bigl(d_x^{j_1}f\bigr)(x' - x) \bigl(d_x^{j_2}p\bigr)(x' - x) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \nonumber \\
+& = \sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1} \frac{\varepsilon^{j_1 + j_2 - 2}}{j_1!j_2!}  \underbrace{\int_{B(0,1)} d_x^{j_1}f(z) d_x^{j_2}p(z) \eta(\|z\|) \,dz}_{:= g_{j_1,j_2}(x)} \label{pf:approximation_error_nonlocal_laplacian_3}
+\end{align}
+The nature of $g_{j_1,j_2}(x)$ depends on the sum $j_1 + j_2$.
+Since $d_x^{j_1}f d_x^{j_2}$ is an order $j_1 + j_2$ (multivariate) monomial, we have (see Section~\ref{subsec:taylor_expansion}) that whenever $j_1 + j_2$ is odd,
+\begin{equation*}
+g_{j_1,j_2}(x) = \int_{\mc{X}} d_x^{j_1}f(z) d_x^{j_2}p(z) \eta(\|z\|) \,dz = 0.
+\end{equation*}
+In particular this is the case when $j_1 = 1$ and $j_2 = 0$, which means that if $s = 2$, then $G_1(x) = 0$.
+On the other hand if $s \geq 3$, then the lowest order term in~\eqref{pf:approximation_error_nonlocal_laplacian_3} is when $j_1 + j_2 = 2$, so that either $j_1 = 1$ and $j_2 = 1$, or $j_1 = 2$ and $j_2 = 0$.
+It follows that
+\begin{align*}
+g_{1,1}(x) + \frac{1}{2}g_{2,0}(x) & = \int_{\mc{X}} d_x^{1}f(z) d_x^{1}p(z) \eta(\|z\|) \,dz + \frac{p(x)}{2} \int_{\mc{X}} d_x^{2}f(z) \eta(\|z\|) \,dz \\
+& = \sum_{i_1 = 1}^{d} \sum_{i_2 = 1}^{d} D^{e_{i_1}}f(x) D^{e_{i_2}}p(x) \int_{\mc{X}} z^{e_{i_1} + e_{i_2}} \eta(\|z\|) \,dz + \frac{p(x)}{2} \sum_{i_1 = 1}^{d} \sum_{i_2 = 1}^{d}D^{e_{i_2}+e_{i_2}}f(x)\int_{\mc{X}} z^{e_{i_1} + e_{i_2}} \eta(\|z\|) \,dz\\
+& = \sum_{i = 1}^{d} D^{e_{i}}f(x) D^{e_{i}}p(x) \int_{\mc{X}} z^2 \eta(\|z\|) \,dz + \frac{p(x)}{2} \sum_{i = 1}^{d} D^{2e_{i}}f(x)\int_{\mc{X}} z^2 \eta(\|z\|) \,dz\\ 
+& = \sigma_{\eta}\Delta_Pf(x).
+\end{align*}
+Now it remains only to deal with the higher-order terms, where $j_1 + j_2 > 2$, and where it suffices to show that each function $g_{j_1,j_2}$ satisfies~\eqref{pf:approximation_error_nonlocal_laplacian_1.5} for $j = \min\{j_1 + j_2 - 2,s - 2\}$.
+Now it is helpful to write $g_{j_1,j_2}$ using multi-index notation,
+\begin{align*}
+g_{j_1,j_2}(x) = \sum_{|\alpha_1| = j_1} \sum_{|\alpha_2| = j_2} D^{\alpha_1}f(x) D^{\alpha_2}p(x) \int_{B(0,1)} z^{\alpha_1 + \alpha_2} \eta(\|z\|) \,dz,
+\end{align*}
+where we note that $|\int_{B(0,1)} z^{\alpha_1 + \alpha_2} \eta(\|z\|) \,dz| < \infty$ for all $\alpha_1, \alpha_2$, by the assumption that $\eta$ is Lipschitz on its support.
+Finally, by H\"{o}lder's inequality we have that
+\begin{align*}
+\|D^{\alpha_1}f D^{\alpha_2}p\|_{H^{s - (j + 2)}(\mc{X})} & \leq \|D^{\alpha_1}f\|_{H^{s - (j + 2)}(\mc{X})} \|D^{\alpha_2}p\|_{C^{s - (j + 2)}(\mc{X})} \\
+& \leq \|D^{\alpha_1}f\|_{H^{s - j_1}(\mc{X})} \|D^{\alpha_2}p\|_{C^{s - (j_2 + 1)}(\mc{X})} \\
+& \leq M \cdot \|p\|_{C^{s - 1}(\mc{X})},
+\end{align*}
+and summing over all $|\alpha_1| = j_1$ and $|\alpha_2| = j_2$ establishes that $g_{j_1,j_2}$ satisfies~\eqref{pf:approximation_error_nonlocal_laplacian_1.5}.
+
+\emph{Estimate on $G_2(x)$.}
+Note immediately that $G_2(x) = 0$ if $s = 1$.
+Otherwise if $s \geq 2$, then $q = s - 1$.Recalling that $|r_{x + z\varepsilon}^{s - 1}(x; p)| \leq C\varepsilon^{s - 1}\|p\|_{C^{s - 1}(\mc{X})}$ for any $z \in B(0,1)$, and that $d_x^jf(\cdot)$ is a $j$-homogeneous function, we have that
+\begin{align}
+|G_2(x)| & \leq \frac{1}{\varepsilon^{d + 2}}\sum_{j = 1}^{s - 1} \frac{1}{j!}\int_{\mc{X}} \Bigl|\bigl(d_x^{j}f\bigr)(x' - x)\Bigr| \cdot \Bigl|r_{x'}^{s - 1}(x;p)\Bigr| \cdot \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \nonumber \\
+& = \frac{1}{\varepsilon^2} \sum_{j = 1}^{s - 1} \frac{1}{j!}\int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(\varepsilon z)\Bigr| \cdot |r_{x + z\varepsilon}^{s - 1}(x;p)| \cdot \eta(\|z\|) \,dz \nonumber \\
+& \leq C\varepsilon^{s - 2}\|p\|_{C^{s - 1}(\mc{X})} \sum_{j = 1}^{s - 1} \frac{1}{j!} \int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(z)\Bigr| \cdot \eta(\|z\|) \,dz \label{pf:approximation_error_nonlocal_laplacian_4}.
+\end{align}
+Furthermore, for each $j = 1,\ldots,s - 1$ convolution of $d_x^jf$ with $\eta$ only decreases the $\Leb^2(\mc{X}_{\varepsilon})$ norm, meaning
+\begin{equation}
+\label{pf:approximation_error_nonlocal_laplacian_5}
+\begin{aligned}
+\int_{\mc{X}_{\varepsilon}} \biggl(\int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(z)\Bigr| \cdot \eta(\|z\|) \,dz\biggr)^2 \,dx & \leq \int_{\mc{X}_{\varepsilon}} \biggl(\int_{B(0,1)} \Bigl|\bigl(d_x^jf\bigr)(z)\Bigr|^2 \eta(\|z\|)\,dz \biggr) \cdot \biggl(\int_{B(0,1)} \eta(\|z\|) \,dz \biggr) \,dx \\
+& \leq \int_{B(0,1)} \int_{\mc{X}_{\varepsilon}} \Bigl[\bigl(d^jf\bigr)(x)\Bigr]^2 \eta(\|z\|) \,dx  \,dz \\
+& \leq \|d^jf\|_{\Leb^2(\mc{X_{\varepsilon}})}^2.
+\end{aligned}
+\end{equation}
+In the above, we have used that $|d_x^jf(z)| \leq |d^jf(x)|$ for all $z \in B(0,1)$, and the normalization $\int \eta(\|z\|) \,dz = 1$. 
+Combining this with~\eqref{pf:approximation_error_nonlocal_laplacian_4}, we conclude that
+\begin{align*}
+\int_{\mc{X}_{\varepsilon}} |G_2(x)|^2 \,dx & \leq C \Bigl(\varepsilon^{s - 2}\|p\|_{C^{s - 1}(\mc{X})}\Bigr)^2 \sum_{j = 1}^{s - 1} \int_{\mc{X}_{\varepsilon}}\biggl(\frac{1}{j!} \int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(z)\Bigr| \cdot \Bigl|\eta(\|z\|)\Bigr| \,dz\biggr)^2 \,dx \\
+& \leq C \Bigl(\varepsilon^{s - 2}\|p\|_{C^{s - 1}(\mc{X})} \sum_{j = 1}^{s - 1}\Bigr)^2 \|d^ju\|_{\Leb^2(\mc{X_{\varepsilon}})}^2,
+\end{align*}
+establishing the desired estimate.
+
+\emph{Estimate on $G_3(x)$.}
+We begin with a pointwise upper bound on $|J_3(x)|$, 
+\begin{align*}
+|G_3(x)| & \leq \frac{p_{\max}}{\varepsilon^{2 + d}} \int_{\mc{X}} \bigl|r_{x'}^s(x;f)\bigr| \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \\
+& = \frac{p_{\max}}{\varepsilon^2} \int_{B(0,1)} \bigl|r_{x + \varepsilon z}^s(x;f)\bigr| \eta(\|z\|) \,dz,
+\end{align*}
+and applying the Cauchy-Schwarz inequality, we deduce a pointwise upper bound on $|G_3(x)|^2$,
+\begin{align*}
+|G_3(x)|^2 & \leq \biggl(\frac{p_{\max}}{\varepsilon^2}\biggr)^2 \cdot \biggl(\int_{B(0,1)} \bigl|r_{x + \varepsilon z}^s(x;u)\bigr|^2 \eta(\|z\|)\,dz\biggr) \cdot \biggl(\int_{B(0,1)} \eta(\|z\|) \,dz\biggr) \\
+& \leq \biggl(\frac{p_{\max}}{\varepsilon^2}\biggr)^2 \int_{B(0,1)} \bigl|r_{x + \varepsilon z}^s(x;u)\bigr|^2 \eta(\|z\|) \,dz.
+\end{align*}
+Applying this pointwise over all $x \in \mc{X}_{\varepsilon}$ and integrating, we obtain
+\begin{align*}
+\int_{\mc{X}_{\varepsilon}} |G_3(x)|^2 \,dx & \leq \biggl(\frac{p_{\max}}{\varepsilon^2}\biggr)^2 \int_{\mc{X}_{\varepsilon}} \int_{B(0,1)} \bigl|r_{x + \varepsilon z}^s(x;f)\bigr|^2 \eta(\|z\|) \,dz \,dx \\
+& = \biggl(\frac{p_{\max}}{\varepsilon^2}\biggr)^2 \int_{B(0,1)} \int_{\mc{X}_{\varepsilon}} \bigl|r_{x + \varepsilon z}^s(x;f)\bigr|^2 \eta(\|z\|) \,dx \,dz \\
+& \leq \biggl(\frac{p_{\max}\varepsilon^s}{\varepsilon^2}\biggr)^2  \|d^sf\|_{L^2(\mc{X}_{\varepsilon})}^2,
+\end{align*}
+with the last inequality following from~\eqref{eqn:sobolev_remainder_term}.
+Noting that $p_{\max} = \|p\|_{C^0(\mc{X})} \leq \|p\|_{C^{s - 1}(\mc{X})}$, we see that this is a sufficient bound on $\|G_3\|_{\Leb^2(\mc{X}_{\varepsilon})}$.
+
+\underline{\textit{Proof of~\eqref{pf:approximation_error_nonlocal_laplacian_0} and~\eqref{pf:approximation_error_nonlocal_laplacian_1}, induction step.}}
+We now assume that~\eqref{pf:approximation_error_nonlocal_laplacian_0} and~\eqref{pf:approximation_error_nonlocal_laplacian_1} hold for all order up to some $k$, and show that they then hold for order $k + 1$ as well.
+The proof is relatively straightforward, once we introduce a bit of notation.
+Namely, for any $\ell,j \in \mathbb{N}$ such that $1 \leq j \leq \ell \leq$, we will use $g_j^{\ell}$ to refer to a function satisfying
+\begin{equation}
+\label{pf:approximation_error_nonlocal_laplacian_6}
+\|g_j^{\ell}\|_{H^{\ell - j}(\mc{X}_{(k + 1)\varepsilon})} \leq C \|p\|_{C^{q}(\mc{X})}^{k + 1} M.
+\end{equation}
+Note that $g_j^{\ell}(x) = g_{(s - \ell) + j}(x)$, so that $g_j^{s}(x) = g_j(x)$.
+As before, the functions $g_j^{\ell}$ may change from line to line, but will always satisfy~\eqref{pf:approximation_error_nonlocal_laplacian_6}.
+We immediately illustrate the purpose of this notation.
+Suppose $g \in H^{\ell}(\mc{X}_{k\varepsilon}; C \|p\|_{C^{q}(\mc{X})}^k M)$ for some $\ell \leq s$.
+If $\ell \leq 2$, then by the inductive hypothesis, it follows that for any $x \in \mc{X}_{(k + 1)\varepsilon}$
+\begin{equation}
+\label{pf:approximation_error_nonlocal_laplacian_7}
+L_{P,\varepsilon}g(x) = g_{\ell}^{\ell}(x) \varepsilon^{\ell - 2}.
+\end{equation} 
+On the other hand if $2 < \ell \leq s$, then by the inductive hypothesis, it follows that for any $x \in \mc{X}_{(k + 1)\varepsilon}$,
+\begin{equation}
+\label{pf:approximation_error_nonlocal_laplacian_8}
+L_{P,\varepsilon}g(x) = \sigma_{\eta} \Delta_Pg(x) + \sum_{j = 1}^{\floor{(\ell - 1)/2} - 1} g_{2j + 2}^{\ell}(x) \varepsilon^{2j} + g_{\ell}^{\ell}(x) \varepsilon^{\ell - 2}.
+\end{equation}
+
+\emph{Proof of \eqref{pf:approximation_error_nonlocal_laplacian_0}.} If $s \leq 2(k + 1)$, then by the inductive hypothesis it follows that for all $x \in \mc{X}_{k\varepsilon}$, we have $L_{P,\varepsilon}^kf(x) = g_{s}(x) \cdot \varepsilon^{s - 2k}$, for some $g_s \in L^2(\mc{X}_{k\varepsilon}, C\|p\|_{C^{s - 1}(\mc{X})}^k M)$.
+Note that we may know more about $L_P^kf(x)$ than simply that it is bounded in $L^2$-norm, but a bound in $L^2$-norm suffices.
+In particular, from such a bound along with~\eqref{pf:approximation_error_nonlocal_laplacian_7} we deduce that for any $x \in \mc{X}_{(k + 1)\varepsilon}$,
+\begin{equation}
+\label{pf:approximation_error_nonlocal_laplacian_8.5}
+L_{P,\varepsilon}^{k + 1}f(x) = \bigl(L_{P,\varepsilon} \circ L_{P,\varepsilon}^k f)(x)= L_{P,\varepsilon} g_s(x)\varepsilon^{s - 2k} = g_{s}^{s}(x) \varepsilon^{s - 2(k + 1)},
+\end{equation}
+establishing~\eqref{pf:approximation_error_nonlocal_laplacian_0}.
+
+\emph{Proof of \eqref{pf:approximation_error_nonlocal_laplacian_1}.} If $s > 2(k + 1)$, then by the inductive hypothesis we have that for all $x \in \mc{X}_{k\varepsilon}$, 
+\begin{equation*}
+L_{P,\varepsilon}^kf(x) = \sigma_{\eta}^k \Delta_P^kf(x) + \sum_{j = 1}^{\floor{(s - 1)/2} - k} g_{2(j + k)}(x) \varepsilon^{2j} + g_s(x) \varepsilon^{s - 2k}.
+\end{equation*}
+Thus for any $x \in \mc{X}_{(k + 1)\varepsilon}$, 
+\begin{equation*}
+L_{P,\varepsilon}^{k + 1}f(x) = \bigl(L_{P,\varepsilon} \circ L_{P,\varepsilon}^k f\bigr)(x) = \sigma_{\eta}^k L_{P,\varepsilon}\Delta_P^kf(x) + \sum_{j = 1}^{\floor{(s - 1)/2} - k} L_{P,\varepsilon}g_{2(j + k)}(x) \varepsilon^{2j} + L_{P,\varepsilon}g_s(x) \varepsilon^{s - 2k}
+\end{equation*}
+There are three terms on the right hand side of this equality, and we now analyze each separately.
+\begin{enumerate}
+	\item Noting that $\Delta_P^kf \in H^{s - 2k}(\mc{X}; C\|p\|_{C^{s - 1}(\mc{X})}^kM)$, we use~\eqref{pf:approximation_error_nonlocal_laplacian_8} to derive that
+	\begin{align}
+	L_{P,\varepsilon}\Delta_P^kf(x) & = \sigma_{\eta} \Delta_P^{k + 1}f(x) + \sum_{j = 1}^{(s - 2k - 1)/2 - } g_{2j + 2}^{s - 2k}(x)\varepsilon^{2j} + g_{s - 2k}^{s - 2k}(x) \varepsilon^{s - 2k - 2} \nonumber \\
+	& = \sigma_{\eta} \Delta_P^{k + 1}f(x) + \sum_{j = 1}^{(s - 1)/2 - (k + 1)} g_{2(k + 1 + j)}(x)\varepsilon^{2j} + g_{s}(x) \varepsilon^{s - 2(k + 1)}, \label{pf:approximation_error_nonlocal_laplacian_9}
+	\end{align}
+	where in the second equality we have simply used the fact $g_j^{\ell}(x) = g_{(s - \ell) + j}(x)$ to rewrite the equation.
+	\item Suppose $j < \floor{(s - 1)/2} - k$. Then we use~\eqref{pf:approximation_error_nonlocal_laplacian_8} to derive that
+	\begin{align*}
+	L_{P,\varepsilon}g_{2(j + k)}(x) & = \sigma_{\eta}\Delta_P g_{2(j + k)}(x) + \sum_{i = 1}^{\floor{(s - 2j - 2k - 1)/2} - 1} g_{2(i + 1)}^{s - 2(j + k)}(x)\varepsilon^{2i} + g_{s - 2(j + k)}^{s - 2(j + k)}(x) \varepsilon^{s - 2(j + k + 1)} \\
+	& = g_{2(j + k + 1)}(x) + \sum_{i = 1}^{\floor{(s - 1)/2} - (j + k + 1)} g_{2(i + j + k + 1)}(x)\varepsilon^{2i} + g_{s}(x) \varepsilon^{s - 2(j + k + 1)},
+	\end{align*}
+	where in the second equality we have again used $g_j^{\ell}(x) = g_{(s - \ell) + j}(x)$, and also written $\sigma_{\eta} \Delta_Pf = g_{2}^{s - 2(j + k)} = g_{2(j + k + 1)}$, since the particular dependence on the Laplacian $\Delta_P$ will not matter.
+From here, multiplying by $\varepsilon^{2j}$, we conclude that
+	\begin{align}
+	\varepsilon^{2j} L_{P,\varepsilon}g_{2(j + k)}(x) & = g_{2(j + k + 1)}(x) \varepsilon^{2j} + \sum_{i = 1}^{\floor{(s - 1)/2} - (j + k + 1)} g_{2(i + j + k + 1)}(x)\varepsilon^{2(i + j)} + g_{s}(x) \varepsilon^{s - 2(k + 1)} \nonumber \\
+	& = g_{2(j + k + 1)}(x) \varepsilon^{2j} + \sum_{m = 1}^{\floor{(s - 1)/2} - (k + 1)}  g_{2(m + k + 1)}(x)\varepsilon^{2m} + g_{s}(x) \varepsilon^{s - 2(k + 1)} \label{pf:approximation_error_nonlocal_laplacian_10},
+	\end{align}
+	with the second equality following upon changing variables to $m = i + j$.
+
+	On the other hand if $j = \floor{(s - 1)/2} - k$, then the calculation is much simpler,
+	\begin{equation}
+	\label{pf:approximation_error_nonlocal_laplacian_11}
+	\varepsilon^{2j} L_{P,\varepsilon}g_{2(j + k)}(x) = g_{s - 2(j + k)}^{s - 2(j + k)}(x) \varepsilon^{2j} \varepsilon^{s - 2(j + k) - 2} = g_s(x) \varepsilon^{s - 2(k + 1)}.
+	\end{equation}
+	\item Finally, it follows immediately from~\eqref{pf:approximation_error_nonlocal_laplacian_8} that
+	\begin{equation}
+	\label{pf:approximation_error_nonlocal_laplacian_12}
+	L_{P,\varepsilon}g_s(x) \varepsilon^{s - 2k} = g_{s}(x) \varepsilon^{s - 2(k + 1)}.
+	\end{equation}
+\end{enumerate}
+Plugging~\eqref{pf:approximation_error_nonlocal_laplacian_9}-\eqref{pf:approximation_error_nonlocal_laplacian_12} back into~\eqref{pf:approximation_error_nonlocal_laplacian_8.5} proves the claim.
+
+\subsection{Boundary behavior of non-local Laplacian}
+\label{subsec:boundary_behavior_nonlocal_laplacian}
+
+\subsection{Estimate of non-local Sobolev seminorm}
+\label{subsec:estimate_nonlocal_seminorm}
+
+\subsection{Assorted integrals}
+\label{subsec:integrals}
+
+\begin{lemma}
+	\label{lem:l2estimate_nonlocal_laplacian}
+	Assume Model~\ref{def:model_flat_euclidean}.
+Suppose $f \in L^2(\mc{X};M)$, and let $L_{P,\varepsilon}$ be defined with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_flat_euclidean}.
+Then there exists a constant $C$ which does not depend on $f$ such that
+	\begin{equation}
+	\label{eqn:l2estimate_nonlocal_laplacian}
+	\|L_{P,\varepsilon}f\|_{\Leb^2(\mc{X})} \leq C p_{\max} M \varepsilon^{-2}
+	\end{equation}
+\end{lemma}
+\paragraph{Proof (of Lemma~\ref{lem:l2estimate_nonlocal_laplacian}).}
+We fix a version of $f \in \Leb^2(\mc{X})$, so that we may speak of its pointwise values.
+
+At a given point $x \in \mc{X}$, we can upper bound $|L_{P,\varepsilon}f(x)|^2$ using the Cauchy-Schwarz inequality as follows,
+\begin{align*}
+|L_{P,\varepsilon}f(x)|^2 & \leq \biggl(\frac{p_{\max}}{\varepsilon^{2 + d}}\biggr)^2 \Biggl(\int_\mc{X} \bigl(|f(x')| + |f(x)|\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx'\Biggr)^2 \\
+& \leq \biggl(\frac{p_{\max}}{\varepsilon^{2 + d}}\biggr)^2 \Biggl(\int_\mc{X} \bigl(|f(x')| + |f(x)|\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \cdot \int \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \Biggr) \\
+& = \frac{p_{\max}^2}{\varepsilon^{4 + d}} \int_\mc{X} \bigl(|f(x')| + |f(x)|\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx'.
+\end{align*}
+The equality follows by the assumption $\int_{\Rd} \eta(\|z\|)\,dx = 1$ in~\ref{asmp:kernel_flat_euclidean}. Integrating over all $x \in \mc{X}$, it follows from the triangle inequality that
+\begin{align}
+\|L_{P,\varepsilon}\|_{\Leb^2(\mc{X})}^2 & \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{\mc{X}} \int_{\mc{X}} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx \nonumber \\
+& \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{\mc{X}} \int_{\mc{X}} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx.
+\label{pf:approximation_error_nonlocal_laplacian_2}
+\end{align}
+Finally, using Fubini's Theorem we determine that
+\begin{equation*}
+\int_{\mc{X}} \int{\mc{X}} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx = 2 \int_{\mc{X}} \int_{\mc{X}} |f(x)|^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx \leq 2 \varepsilon^d \int_{\mc{X}} |f(x)|^2 \,dx = 2\varepsilon^d \|f\|_{\Leb^2(\mc{X})}^2
+\end{equation*}
+and in combination with \eqref{pf:approximation_error_nonlocal_laplacian_2} we conclude that
+\begin{equation*}
+\|L_{P,\varepsilon}\|_{\Leb^2(\mc{X})}^2 \leq \frac{4p_{\max}^2}{\varepsilon^4} \|f\|_{\Leb^2(\mc{X})}^2. 
+\end{equation*}
 
@@
 \section{Miscellaneous}
-Here we give some helpful Lemmas used at various points in the above proofs. We also prove that spectral sparsification does not change the rate of convergence of $\wh{f}$.
+Here we give some helpful Lemmas used at various points in the above proofs.
+We also prove that spectral sparsification does not change the rate of convergence of $\wh{f}$.
+Finally, we review notation and relevant facts regarding Taylor expansion.
+
+\subsection{Taylor expansion}
+\label{subsec:taylor_expansion}
+Let $u$ be a function which is $s$ times continuously differentiable at all $x \in \mc{X}$, for $k \in \mathbb{N}\setminus\{0\}$.
+Suppose that for some $h > 0$, $x \in \mc{X}_{h}$ and $x' \in B(x,h)$.
+We write the order-$s$ Taylor expansion of $u(x')$ around $x' = x$ as
+\begin{equation*}
+u(x') = u(x) + \sum_{j = 1}^{s - 1} \frac{1}{j!}\bigl(d_x^{j}u\bigr)(x' - x) + r_{x'}^{s}(x;u)
+\end{equation*}
+For notational convenience we have adopted the convention that $\sum_{j = 1}^{0} a_j = 0$.
+For a given $z \in \Rd$ and $j$-times differentiable function $f: \mc{X} \to \Reals$, we have used the notation $\bigl(d_x^jf\bigr)(z) := \sum_{\abs{\alpha} = j} D^{\alpha}f(x) z^{\alpha}$.
+Thus $\bigl(d_x^{j}f\bigr)(z)$ is a degree-$j$ polynomial---and so a $j$-homogeneous function---in $z$, meaning for any $t \in \Reals$,
+\begin{equation*}
+\bigl(d_x^{j}f\bigr)(tz) = t^{j} \cdot \bigl(d_x^{j}f\bigr)(z).
+\end{equation*}
+The remainder term $r_{x'}$ is given by
+\begin{equation*}
+r_{x'}^s(x;f) = \frac{1}{(j - 1)!} \int_{0}^{1}(1 - t)^{j - 1} \bigl(d_{x + t(x' - x)}^{s}f\bigr)(x' - x) \,dt,
+\end{equation*}
+where we point out that the integral makes sense because $x + t(x' - x) \in B(x,h) \subseteq \mc{X}$.
+We now give estimates on the remainder term in both sup-norm and $L^2(\mc{X}_{h})$ norm, each of which hold for any $z \in B(0,1)$.
+In sup-norm, we have that
+\begin{equation*}
+\sup_{x \in \mc{X}_{h}}|r_{x + hz}^j(x;f)| \leq C h^j \|f\|_{C^j(\mc{X})},
+\end{equation*}
+whereas in $L^2(\mc{X}_{h})$ norm we have,
+\begin{equation}
+\label{eqn:sobolev_remainder_term}
+\int_{\mc{X}_{h}} \bigl|r_{x + thz}^j(x;f)\bigr|^2 \,dx \leq h^{2j} \int_{\mc{X}_{h}} \int_{0}^{1} |d_{x + thz}^jf(z)|^2 \,dt \,dx \leq h^{2j} \|d^jf\|_{\Leb^2(\mc{X})}^2.
+\end{equation}
+
+Finally, we recall some facts regarding the interaction between smoothing kernels and polynomials.
+Let $q_j(z)$ be an arbitrary degree-$j$ (multivariate) polynomial.
+If $\eta$ is a radially symmetric kernel and $j$ is odd, then by symmetry it follows that
+\begin{equation*}
+\int_{B(0,1)} q_j(z) \eta(\|z\|) \,dz = 0.
+\end{equation*}
+On the other hand, if $\psi$ is an order-$s$ kernel for some $s > j$, then by converting to polar coordinates we can verify that
+\begin{equation*}
+\int_{B(0,1)} q_j(z) \eta(\|z\|) \,dz = 0.
+\end{equation*}
+
+\textcolor{red}{(TODO): Check with Ryan and Siva that (i) they believe these facts, and (ii) these facts do not require additional justification.}

### 2021-04-02 manuscripts/laplacian_eigenmaps/appendix.tex
Completed analysis of non-local Sobolev seminorm in flat Euclidean case.

@@
 	\label{pf:graph_seminorm_ho_1}
-	\dotp{L_{n,\varepsilon}^sf}{f}_n \leq \frac{1}{\delta} \dotp{L_{P,\varepsilon}^sf}{f}_{P} + \frac{C\varepsilon^2}{n\varepsilon^{2 + d}}M^2.
+	\dotp{L_{n,\varepsilon}^sf}{f}_n \leq \frac{1}{\delta} \dotp{L_{P,\varepsilon}^sf}{f}_{P} + \frac{C\varepsilon^2}{\delta n\varepsilon^{2 + d}}M^2.
 	\end{equation}
-	with probability at least $1 - \delta$.
+	with probability at least $1 - 2\delta$.
 
 	We term the first term on the right hand side the \emph{non-local Sobolev semi-norm}, as it is a kernelized approximation to the Sobolev semi-norm $\dotp{\Delta_P^sf}{f}_{P}$.
-The second term on the right hand side is a pure bias term, which as we will see is negligible compared to the non-local quadratic form as long as $\varepsilon \ll n^{-1/(2(s -1 + d))}$.
+The second term on the right hand side is a pure bias term, which as we will see is negligible compared to the non-local Sobolev semi-norm as long as $\varepsilon \ll n^{-1/(2(s -1 + d))}$.
 	\item~In Section~\ref{subsec:approximation_error_nonlocal_laplacian}, we show that when $x$ is sufficiently in the interior of $\mc{X}$, then $L_{P,\varepsilon}^kf(x)$ is a good approximation to $\Delta_P^kf(x)$, as long as $f \in H^{s}(\mc{X})$ and $p \in C^{s - 1}(\mc{X})$ for some $s \geq 2k + 1$.
-	\item~In Section~\ref{subsec:boundary_behavior_nonlocal_laplacian}, we show that when $x$ is sufficiently near the boundary of $\mc{X}$, then $L_{P,\varepsilon}^kf(x)$ is close to $0$, as long as $f \in H_0^{s}(\mc{X})$ for some $s \geq 2k$.
+	\item~In Section~\ref{subsec:boundary_behavior_nonlocal_laplacian}, we show that when $x$ is sufficiently near the boundary of $\mc{X}$, then $L_{P,\varepsilon}^kf(x)$ is close to $0$, as long as $f \in H_0^{s}(\mc{X})$ for some $s > 2k$.
 	\item~In Section~\ref{subsec:estimate_nonlocal_seminorm}, we use the results of the preceding two sections to show that if $f \in H_0^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$, there exists a constant $C$ which does not depend on $f$ such that
@@
 \paragraph{Proof (of Proposition~\ref{prop:graph_seminorm_ho}).}
-Proposition~\ref{prop:graph_seminorm_ho} follows immediately from~\eqref{pf:graph_seminorm_ho_1} and~\eqref{pf:graph_seminorm_ho_2}.
+Proposition~\ref{prop:graph_seminorm_ho} follows immediately from~\eqref{pf:graph_seminorm_ho_1} and~\eqref{pf:graph_seminorm_ho_2}. \qed
+
+One note regarding notation: suppose a function $g \in H^{\ell}(U)$, where $\ell \in \mathbb{N}$ and $U$ is an open set.
+Let $V$ be another open set, compactly contained within $U$.
+Then we will use the notation $g \in H^{\ell}(V)$ to mean that the restriction $\restr{g}{V}$ of $g$ to $V$ belongs to $H^{\ell}(V)$.
 
@@
 
+In Lemma~\ref{lem:approximation_error_nonlocal_laplacian_boundary}, we establish that if $f$ is Sobolev smooth of order $s > 2k$ and zero-trace, then near the boundary of $\mc{X}$ the non-local Laplacian $L_{P,\varepsilon}^kf$ is close to $0$ in the $\Leb^2$-sense.
+\begin{lemma}
+	\label{lem:approximation_error_nonlocal_laplacian_boundary}
+	Assume Model~\ref{def:model_flat_euclidean}.
+Let $s,k \in \mathbb{N}$.
+Suppose that $f \in H_0^{s}(\mc{X};M)$.
+Then there exist numbers $c,C > 0$ that do not depend on $M$, such that for all $\varepsilon < c$,
+	\begin{equation*}
+	\|L_{P,\varepsilon}^kf\|_{L^2(\partial_{k\varepsilon}\mc{X})}^2 \leq C \varepsilon^{2(s - 2k)}M^2.
+	\end{equation*}
+\end{lemma}
+
+\paragraph{Proof (of Lemma~\ref{lem:approximation_error_nonlocal_laplacian_boundary})}
+Applying Lemma~\ref{lem:l2estimate_nonlocal_laplacian}, we have that
+\begin{equation*}
+\|L_{P,\varepsilon}^kf\|_{L^2(\partial_{k\varepsilon}(\mc{X}))}^2 \leq \frac{(Cp_{\max})^{2}}{\varepsilon^4} \|L_{P,\varepsilon}^{k - 1}f\|_{L^2(\partial_{k\varepsilon}(\mc{X}))}^2 \leq \cdots \leq \frac{(Cp_{\max})^{2}}{\varepsilon^{4k}} \|f\|_{L^2(\partial_{k\varepsilon}(\mc{X}))}^2
+\end{equation*}
+Thus it remains to show that for all $\varepsilon < c$,
+\begin{equation}
+\label{pf:approximation_error_nonlocal_laplacian_boundary_0}
+\|f\|_{L^2(\partial_{k\varepsilon}(\mc{X}))}^2 = \int_{\partial_{k\varepsilon}(\mc{X})} \bigl(f(x)\bigr)^2 \,dx \leq C_1 \varepsilon^{2s} \|f\|_{H^s(\mc{X})}^2.
+\end{equation}
+We will build to~\eqref{pf:approximation_error_nonlocal_laplacian_boundary_0} by a series of intermediate steps, following the same rough structure as the proof of Theorem 18.1 in \citet{leoni2017}.
+For simplicity, we will take $k = 1$; the exact same proof applies to the general case upon assuming $\varepsilon < c/k$.
+
+\underline{\textit{Step 1: Local Patch.}}
+To begin, we assume that for some $c_0 > 0$ and a Lipschitz mapping $\phi: \Reals^{d - 1} \to [-c_0,c_0]$, we have that $f \in C_c^{\infty}(U_{\phi}(c_0))$, where 
+\begin{equation*}
+U_{\phi}(c_0) = \Bigl\{y \in Q(0,c_0): \phi(y_{-d}) \leq y_d\Bigr\}, 
+\end{equation*}
+and here $Q(0,c_0)$ is the $d$-dimensional cube of side length $c_0$, centered at $0$.
+We will show that for all $0 < \varepsilon < c_0$, and for the tubular neighborhood $V_{\phi}(\varepsilon) = \{y \in Q(0,c_0): \phi(y_{-d}) \leq y_d \leq \phi(y_{-d}) + \varepsilon\}$, we have that
+\begin{equation*}
+\int_{V_{\phi}(\varepsilon)} |f(x)|^2 \,dx \leq C\varepsilon^{2s} \|f\|_{H^s(U_{\phi}(c_0))}^2.
+\end{equation*}
+For a given $y = (y',y_d) \in V_{\phi}(\varepsilon)$, let $y_0 = (y',\phi(y'))$.
+Taking the Taylor expansion of $f(y)$ around $y = y_0$ because $u$ is compactly supported in $V_{\phi}$ it follows that,
+\begin{align*}
+f(y) & = f(y_0) + \sum_{j = 1}^{s - 1} \frac{1}{j!} D^{je_d}f(y_0) \bigl(y_d - \phi(y')\bigr)^j + \frac{1}{(s - 1)!}\int_{\phi(y')}^{y_d} (1 - t)^{s - 1} D^{se_d}f(y',z) \bigl(y_d - z\bigr)^{s - 1} \,dz \Longrightarrow\\
+|f(y)| & \leq C\varepsilon^{s - 1}\int_{\phi(y')}^{y_d} \bigl|D^{se_d}f(y',z)\bigr| \,dz. 
+\end{align*}
+Consequently, by squaring both sides and applying Cauchy-Schwarz, we have that
+\begin{equation*}
+|f(y)|^2 \leq C\varepsilon^{2(s - 1)} \biggl(\int_{\phi(y')}^{y_d} \bigl|D^{se_d}f(y',z)\bigr| \,dz\biggr)^2 \leq C\varepsilon^{2s - 1} \int_{\phi(y')}^{y_d} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz.
+\end{equation*}
+Applying this bound for each $y \in V_{\phi}(\varepsilon)$, and then integrating, we obtain
+\begin{align}
+\int_{V_{\phi}(\varepsilon)} |f(y)|^2 \,dy & \leq \int_{Q_{d - 1}(c_0)} \int_{\phi(y')}^{\phi(y') + \varepsilon} |f(y',y_d)|^2 \,dy_d \,dy' \nonumber \\
+& \leq C\varepsilon^{2s - 1}\int_{Q_{d - 1}(c_0)}  \int_{\phi(y')}^{\phi(y') + \varepsilon} \int_{\phi(y')}^{y_d} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz \,dy_d \,dy' \label{pf:approximation_error_nonlocal_laplacian_boundary_1}
+\end{align}
+where we have written $Q_{d - 1}(0,c_0)$ for the $d - 1$ dimensional cube of side length $c_0$, centered at $0$. Exchanging the order of the inner two integrals then gives
+\begin{align*}
+\int_{\phi(y')}^{\phi(y') + \varepsilon} \int_{\phi(y')}^{y_d} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz \,dy_d & = \int_{\phi(y')}^{\phi(y') + \varepsilon} \int_{z}^{\varepsilon} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dy_d \,dz \\
+& \leq C \varepsilon \int_{\phi(y')}^{\phi(y') + \varepsilon} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz \\
+& \leq C \varepsilon \int_{\phi(y')}^{c_0} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz.
+\end{align*}
+Finally, plugging back into~\eqref{pf:approximation_error_nonlocal_laplacian_boundary_1}, we conclude that
+\begin{equation*}
+\int_{V_{\phi}(\varepsilon)} |f(y)|^2 \,dy \leq C \varepsilon^{2s} \int_{Q_{d - 1}(0,c_0)} \int_{\phi(y')}^{c_0} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz \,dy' \leq C \varepsilon^{2s} |u|_{H^s(U_{\phi}(c_0))}^2.
+\end{equation*}
+
+\underline{\textit{Step 2: Rigid motion of local patch.}} Now, suppose that at a point $x_0 \in \partial \mc{X}$, there exists a rigid motion $T: \Rd \to \Rd$ for which $T(x_0) = 0$, and a number $C_0$ such that for all $\varepsilon \cdot C_0 \leq c_0$,
+\begin{equation*}
+T\bigl(Q_{T}(x_0,c_0) \cap \partial_{\varepsilon}\mc{X}\bigr) \subseteq V_{\phi}\bigl(C_0\varepsilon\bigr) \quad\textrm{and}\quad T\bigl(Q_T(x_0,c_0) \cap \mc{X}\bigr) = U_{\phi}(c_0).
+\end{equation*}
+Here $Q_{T}(x_0,c_0))$ is a (not necessarily coordinate-axis-aligned) cube of side length $c_0)$, centered at $x_0$.
+Define $v(y) := f(T^{-1}(y))$ for $y \in U_{\phi}(c_0)$.
+If $u \in C_c^{\infty}(\mc{X})$, then $v \in C_c^{\infty}(U_{\phi}(c_0))$, and moreover $\|v\|_{H^s(U_{\phi}(c_0))}^2 = \|f\|_{H^s(Q_{T}(x_0,c_0) \cap \mc{X})}^2$.
+Therefore, using the upper bound that we derived in Step 1,
+\begin{equation*}
+\int_{V_{\phi}(C_0 \cdot \varepsilon)} |v(y)|^2 \,dy \leq C \varepsilon^{2s} \|v\|_{H^s(U_{\phi}(c_0))}^2,
+\end{equation*}
+we conclude that
+\begin{align*}
+\int_{Q_{T}(x_0,c_0) \cap \partial_{\varepsilon}\mc{X}} |f(x)|^2 \,dx & = \int_{T(Q_T(x_0,c_0)) \cap \partial_{\varepsilon}\mc{X})} |v(y)|^2 \,dy \\
+& \leq \int_{V_{\phi}(C_0 \cdot \varepsilon)} |v(y)|^2 \,dy \\
+& \leq C \varepsilon^{2s} \|v\|_{H^s(U_{\phi}(c_0))}^2 = C \varepsilon^{2s} \|f\|_{H^s(Q_{T}(x_0,c_0)) \cap \mc{X})}^2 \leq C \varepsilon^{2s} \|f\|_{H^s(\mc{X})}^2.
+\end{align*}
+
+\underline{\textit{Step 3: Lipschitz domain}}.
+Finally, we deal with the case where $\mc{X}$ is assumed to be an open, bounded subset of $\Rd$, with Lipschitz boundary.
+In this case, at every $x_0 \in \partial \mc{X}$, there exists a rigid motion $T_{x_0}: \Rd \to \Rd$ such that $T_{x_0}(x_0) = 0$, a number $c_0(x_0)$, a Lipschitz function $\phi_{x_0}:\Reals^{d - 1} \to [-c_0,c_0]$, and a number $C_0(x_0)$, such that for all $\varepsilon \cdot C_0(x_0) \leq c_0(x_0)$,
+\begin{equation*}
+T\bigl(Q_{T}(x_0,c_0(x_0)) \cap \partial_{\varepsilon}\mc{X}\bigr) \subseteq V_{\phi}\bigl(C_0(x_0) \cdot \varepsilon\bigr) \quad\textrm{and}\quad T\bigl(Q_T(x_0,c_0(x_0)) \cap \mc{X}\bigr) = U_{\phi}(c_0(x_0)).
+\end{equation*}
+Therefore for every $x_0 \in \partial \mc{X}$, it follows from the previous step that
+\begin{equation*}
+\int_{Q_{T_{x_0}}(x_0,c_0(x_0)) \cap \partial_{\varepsilon}\mc{X}} |f(x)|^2 \,dx \leq C(x_0) \varepsilon^{2s} \|f\|_{H^s(\mc{X})}^2,
+\end{equation*}
+where on the right hand side $C(x_0)$ is a constant that may depend on $x_0$, but not on $u$ or $\varepsilon$.
+
+We conclude by taking a collection of cubes that covers $\partial_{\varepsilon}\mc{X}$ for all $\epsilon$ sufficiently small.
+First, we note that by a compactness argument there exists a finite subset of the collection of cubes $\{Q_{T_{x_0}}(x_0,c_0(x_0)/2): x_0 \in \partial\mc{X} \}$ which covers $\partial \mc{X}$, say $Q_{T_{x_1}}(x_1,c_0(x_1)/2),\ldots, Q_{T_{x_N}}(x_N,c_0(x_N)/2)$.
+Then, for any $\varepsilon \leq \min_{i = 1,\ldots,N} c_0(x_i)/2$, it follows from the triangle inequality that
+\begin{equation*}
+\partial_{\varepsilon}\mc{X} \subseteq \bigcup_{i = 1}^{N} Q_{T_{x_i}}(x_i, c_0(x_i)).
+\end{equation*}
+As a result,
+\begin{equation*}
+\int_{\partial_{\varepsilon}\mc{X}} |f(x)|^2 \leq \sum_{i = 1}^{N} \int_{Q_{T_{x_i}}(x_i, c_0(x_i)) \cap \partial_{\varepsilon}(\mc{X})} |f(x)|^2 \leq  \varepsilon^{2s} \|f\|_{H^s(\mc{X})}^2 \sum_{i = 1}^{N}C_0(x_i),
+\end{equation*}
+which proves the claim of~\eqref{pf:approximation_error_nonlocal_laplacian_boundary_0}.
+
 \subsection{Estimate of non-local Sobolev seminorm}
 \label{subsec:estimate_nonlocal_seminorm}
+
+Now, we use the results of the preceding two sections to prove~\eqref{pf:graph_seminorm_ho_2}.
+We will divide our analysis in two cases, depending on whether $s$ is odd or even, but before we do this we state some facts that will be applicable to both cases.
+First, we  recall that $L_{P,\varepsilon}$ is self-adjoint in $L^2(P)$, meaning $\dotp{L_{P,\varepsilon}f}{g}_{P} = \dotp{f}{L_{P,\varepsilon}g}_{P}$ for all $f, g \in L^2(P)$.
+We will also use the fact that when $s = 1$, the non-local seminorm is a \emph{non-local Dirichlet energy},
+\begin{equation*}
+\dotp{L_{P,\varepsilon}f}{f}_{P} = \frac{1}{\varepsilon^{d + 2}}\int_{\mc{X}} \int_{\mc{X}} \bigl(f(x) - f(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) =: E_{P,\varepsilon}(f;\mc{X}).
+\end{equation*}
+Finally, we recall a result of \textcolor{blue}{(Green 21)}: there exist constants $c_0$ and $C_0$ which do not depend on $M$, such that for all $\varepsilon < c_0$ and for any $f \in H^1(\mc{X};M)$,
+\begin{equation}
+\label{pf:estimate_nonlocal_seminorm_1}
+E_{P,\varepsilon}(f;\mc{X}) \leq C_0 M^2.
+\end{equation}
+\paragraph{Case 1: $s$ odd.}
+Suppose $s$ is odd, so that $s \geq 3$.
+Taking $k = (s - 1)/2$, we use the self-adjointness of $L_{P,\varepsilon}$ to relate the non-local semi-norm $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$ to a non-local Dirichlet energy,
+\begin{equation*}
+\dotp{L_{P,\varepsilon}^sf}{f}_P = \dotp{L_{P,\varepsilon}^{k + 1}f}{L_{P,\varepsilon}^{k}f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}^{k};\mc{X}).
+\end{equation*}
+We now separate this energy into integrals over $\mc{X}_{k\varepsilon}$ and $\partial_{k\varepsilon}(\mc{X})$,
+\begin{align}
+E_{P,\varepsilon}(L_{P,\varepsilon}^{k};\mc{X}) & = \frac{1}{\varepsilon^{d + 2}}\Biggl\{\int_{\mc{X}_{k\varepsilon}} \int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - L_{P,\varepsilon}^kf(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) \nonumber \\
+& \quad + \int_{\partial_{k\varepsilon}\mc{X}} \int_{\partial_{k\varepsilon}\mc{X}} \bigl(L_{P,\varepsilon}^kf(x) - L_{P,\varepsilon}^kf(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x)\Biggr\} \nonumber \\
+& := E_{P,\varepsilon}(L_{P,\varepsilon}^{k};\mc{X}_{k\varepsilon}) + E_{P,\varepsilon}(L_{P,\varepsilon}^{k};\partial_{k\varepsilon}\mc{X}) \label{pf:estimate_nonlocal_seminorm_1.5}
+\end{align}
+and upper bound each energy separately.
+For the first term, we add and substract $\sigma_{\eta}^k\Delta_P^kf(x)$ and $\sigma_{\eta}^k\Delta_P^kf(x')$ within the integrand, then use the triangle inequality and \textcolor{red}{the symmetry between $x$ and $x'$} to deduce that
+\begin{equation}
+\label{pf:estimate_nonlocal_seminorm_2}
+E_{P,\varepsilon}(L_{P,\varepsilon}^{k};\mc{X}_{k\varepsilon}) \leq 3 \sigma_{\eta}^{2k} E_{P,\varepsilon}(\Delta_P^kf;\mc{X}_{k\varepsilon}) + \frac{2}{\varepsilon^{d + 2}}\int_{\mc{X}_{k\varepsilon}} \int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_P^kf(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x).
+\end{equation}
+Noticing that $\Delta_P^kf \in H^1(\mc{X};\|p\|_{C^{s - 1}(\mc{X})}^kM)$, we use~\eqref{pf:estimate_nonlocal_seminorm_1} to conclude that $E_{P,\varepsilon}(\Delta_P^kf;\mc{X}_{k\varepsilon}) \leq C_0M^2$.
+On the other hand, it follows from Assumption~\ref{asmp:kernel_flat_euclidean} and~\eqref{eqn:approximation_error_nonlocal_laplacian_1} that
+\begin{align*}
+\frac{2}{\varepsilon^{d + 2}}\int_{\mc{X}_{k\varepsilon}} \int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_P^kf(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) & \leq \frac{2p_{\max}}{\varepsilon^{2}}\int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_P^kf(x)\bigr)^2 \,dP(x) \\
+& \leq C_1M^2.
+\end{align*}
+Plugging these two bounds into~\eqref{pf:estimate_nonlocal_seminorm_2} gives the desired upper bound on $E_{P,\varepsilon}(L_{P,\varepsilon}^{k};\mc{X}_{k\varepsilon})$.
+
+For the second term in~\eqref{pf:estimate_nonlocal_seminorm_1.5}, we apply Lemmas~\ref{lem:dirichlet_estimate_nonlocal_laplacian} and~\ref{lem:approximation_error_nonlocal_laplacian_boundary} and conclude that,
+\begin{equation*}
+E_{P,\varepsilon}(L_{P,\varepsilon}^{k};\partial_{k\varepsilon}\mc{X}) \leq \frac{4p_{\max}^2}{\varepsilon^2}\|L_{P,\varepsilon}^kf\|_{L^2(\partial_{k\varepsilon}\mc{X})} \leq C M^2.
+\end{equation*}
+\paragraph{Case 2: $s$ even.}
+If $s \in \mathbb{N}$ is even, $s \geq 2$, then letting $k = (s - 2)/2$, the  self-adjointness of $L_{P,\varepsilon}$ implies
+\begin{equation*}
+\dotp{L_{P,\varepsilon}^sf}{f}_P = \|L_{P,\varepsilon}^{k + 1}f\|_P^2.
+\end{equation*}
+As in the first case, we divide the integral up into the interior region $\mc{X}_{k\varepsilon}$ and the boundary region $\partial_{k\varepsilon}\mc{X}$,
+\begin{equation}
+\label{pf:estimate_nonlocal_seminorm_3}
+\|L_{P,\varepsilon}^{k + 1}f\|_P^2 \leq p_{\max} \|L_{P,\varepsilon}^{k + 1}f\|_{\Leb^2(\mc{X})}^2 \leq p_{\max}\biggl\{\int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^{k + 1}f(x)\bigr)^2 \,dP(x) + \int_{\partial_{k\varepsilon}\mc{X}} \bigl(L_{P,\varepsilon}^{k + 1}f(x)\bigr)^2 \,dP(x)\biggr\},
+\end{equation}
+and upper bound each term separately. For the first term, adding and subtracting $\sigma_{\eta}^k \Delta_P^kf(x)$ gives
+\begin{align*}
+\int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^{k + 1}f(x)\bigr)^2 \,dP(x) & \leq 2\int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}\Delta_P^kf(x)\bigr)^2 \,dP(x)  + 2 \int_{\mc{X}_{k\varepsilon}} \Bigl(L_{P,\varepsilon}\bigl(L_{P,\varepsilon}^kf- \sigma_{\eta}\Delta_P^kf\bigr)(x)\Bigr)^2 \,dP(x) \\
+& \overset{(i)}{\leq} CM^2  + 2 \int_{\mc{X}_{k\varepsilon}} \Bigl(L_{P,\varepsilon}\bigl(L_{P,\varepsilon}^kf- \sigma_{\eta}\Delta_P^kf\bigr)(x)\Bigr)^2 \,dP(x) \\
+& \overset{(ii)}{\leq} CM^2  + \frac{Cp_{\max}^2}{\varepsilon^{2}} \|L_{P,\varepsilon}^kf- \sigma_{\eta}\Delta_P^kf\|_{L^2(\mc{X}_{k\varepsilon})}^2 \\
+& \overset{(iii)}{\leq} CM^2,
+\end{align*}
+with $(i)$ following from~\eqref{pf:approximation_error_nonlocal_laplacian_0} since $\Delta_P^kf \in H^2(\mc{X};M\|p\|_{C^{s - 1}(\mc{X})}^l)$, $(ii)$ following from Lemma~\ref{lem:l2estimate_nonlocal_laplacian}, and $(iii)$ following from~\eqref{eqn:approximation_error_nonlocal_laplacian_2}.
+
+Then Lemma~\ref{lem:approximation_error_nonlocal_laplacian_boundary} shows that the second term in~\eqref{pf:estimate_nonlocal_seminorm_3} satisfies
+\begin{equation*}
+\int_{\partial_{k\varepsilon}\mc{X}} \bigl(L_{P,\varepsilon}^{k + 1}f(x)\bigr)^2 \,dP(x) \leq CM^2.
+\end{equation*}
 
@@
 	Assume Model~\ref{def:model_flat_euclidean}.
-Suppose $f \in L^2(\mc{X};M)$, and let $L_{P,\varepsilon}$ be defined with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_flat_euclidean}.
+Suppose $f \in L^2(U;M)$ for an open set $U \subseteq \mc{X}$, and let $L_{P,\varepsilon}$ be defined with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_flat_euclidean}.
 Then there exists a constant $C$ which does not depend on $f$ such that
@@
 	\label{eqn:l2estimate_nonlocal_laplacian}
-	\|L_{P,\varepsilon}f\|_{\Leb^2(\mc{X})} \leq C p_{\max} M \varepsilon^{-2}
+	\|L_{P,\varepsilon}f\|_{\Leb^2(U)} \leq \frac{2 p_{\max}}{\varepsilon^2} \|f\|_{L^2(U)}
+	\end{equation}
+\end{lemma}
+
+\begin{lemma}
+	\label{lem:dirichlet_estimate_nonlocal_laplacian}
+	Assume Model~\ref{def:model_flat_euclidean}.
+Suppose $f \in L^2(U;M)$ for an open set $U \subseteq \mc{X}$, and let $L_{P,\varepsilon}$ be defined with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_flat_euclidean}.
+Then there exists a constant $C$ which does not depend on $f$ such that
+	\begin{equation}
+	\label{eqn:dirichlet_estimate_nonlocal_laplacian}
+	E_{P,\varepsilon}(f;U) \leq \frac{4 p_{\max}^2}{\varepsilon^2} \|f\|_{L^2(U)}^2
 	\end{equation}
@@
 \paragraph{Proof (of Lemma~\ref{lem:l2estimate_nonlocal_laplacian}).}
-We fix a version of $f \in \Leb^2(\mc{X})$, so that we may speak of its pointwise values.
+We fix a version of $f \in \Leb^2(U)$, so that we may speak of its pointwise values.
 
-At a given point $x \in \mc{X}$, we can upper bound $|L_{P,\varepsilon}f(x)|^2$ using the Cauchy-Schwarz inequality as follows,
+At a given point $x \in U$, we can upper bound $|L_{P,\varepsilon}f(x)|^2$ using the Cauchy-Schwarz inequality as follows,
 \begin{align*}
-|L_{P,\varepsilon}f(x)|^2 & \leq \biggl(\frac{p_{\max}}{\varepsilon^{2 + d}}\biggr)^2 \Biggl(\int_\mc{X} \bigl(|f(x')| + |f(x)|\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx'\Biggr)^2 \\
-& \leq \biggl(\frac{p_{\max}}{\varepsilon^{2 + d}}\biggr)^2 \Biggl(\int_\mc{X} \bigl(|f(x')| + |f(x)|\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \cdot \int \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \Biggr) \\
-& = \frac{p_{\max}^2}{\varepsilon^{4 + d}} \int_\mc{X} \bigl(|f(x')| + |f(x)|\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx'.
+|L_{P,\varepsilon}f(x)|^2 & \leq \biggl(\frac{p_{\max}}{\varepsilon^{2 + d}}\biggr)^2 \Biggl(\int_U \bigl(|f(x')| + |f(x)|\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx'\Biggr)^2 \\
+& \leq \biggl(\frac{p_{\max}}{\varepsilon^{2 + d}}\biggr)^2 \Biggl(\int_U \bigl(|f(x')| + |f(x)|\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \cdot \int \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \Biggr) \\
+& = \frac{p_{\max}^2}{\varepsilon^{4 + d}} \int_U \bigl(|f(x')| + |f(x)|\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx'.
 \end{align*}
-The equality follows by the assumption $\int_{\Rd} \eta(\|z\|)\,dx = 1$ in~\ref{asmp:kernel_flat_euclidean}. Integrating over all $x \in \mc{X}$, it follows from the triangle inequality that
+The equality follows by the assumption $\int_{\Rd} \eta(\|z\|)\,dx = 1$ in~\ref{asmp:kernel_flat_euclidean}. Integrating over all $x \in U$, it follows from the triangle inequality that
 \begin{align}
-\|L_{P,\varepsilon}\|_{\Leb^2(\mc{X})}^2 & \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{\mc{X}} \int_{\mc{X}} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx \nonumber \\
-& \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{\mc{X}} \int_{\mc{X}} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx.
+\|L_{P,\varepsilon}\|_{\Leb^2(U)}^2 & \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx \nonumber \\
+& \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx.
 \label{pf:approximation_error_nonlocal_laplacian_2}
@@
 \begin{equation*}
-\int_{\mc{X}} \int{\mc{X}} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx = 2 \int_{\mc{X}} \int_{\mc{X}} |f(x)|^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx \leq 2 \varepsilon^d \int_{\mc{X}} |f(x)|^2 \,dx = 2\varepsilon^d \|f\|_{\Leb^2(\mc{X})}^2
+\int_{U} \int{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx = 2 \int_{U} \int_{U} |f(x)|^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx \leq 2 \varepsilon^d \int_{U} |f(x)|^2 \,dx = 2\varepsilon^d \|f\|_{\Leb^2(U)}^2
 \end{equation*}
@@
 \begin{equation*}
-\|L_{P,\varepsilon}\|_{\Leb^2(\mc{X})}^2 \leq \frac{4p_{\max}^2}{\varepsilon^4} \|f\|_{\Leb^2(\mc{X})}^2. 
+\|L_{P,\varepsilon}\|_{\Leb^2(U)}^2 \leq \frac{4p_{\max}^2}{\varepsilon^4} \|f\|_{\Leb^2(U)}^2. 
 \end{equation*}

### 2021-04-03 manuscripts/laplacian_eigenmaps/appendix.tex
Completed the proof of Proposition 2 in Laplacian Eigenmaps paper.

@@
 \label{subsec:decomposition_graph_seminorm}
-\textcolor{red}{(TODO)}
+
+In Lemma~\textcolor{red}{(?)}, we show that the graph Sobolev semi-norm (V-statistic) decomposes into an unbiased estimate of the non-local Sobolev semi-norm (U-statistic), and a pure bias term.
+We also establish that the pure bias term will be small (in expectation) relative to the U-statistic whenever $\varepsilon$ is sufficiently small.
+\begin{lemma}
+	\label{lem:graph_seminorm_bias}
+	For any $f \in L^2(\mc{X})$, the graph Sobolev semi-norm satisfies
+	\begin{equation}
+	\label{eqn:graph_seminorm_bias_1}
+	\dotp{L_{n,\varepsilon}^sf}{f}_{n} = U_{n,\varepsilon}^{(s)}(f) + B_{n,\varepsilon}^{(s)}(f),
+	\end{equation}
+	such that $\mathbb{E}[U_{n,\varepsilon}^{(s)}(f)] = (n - s - 1)!/n!
+\cdot \dotp{L_{P,\varepsilon}^sf}{f}_P$.
+If additionally $f \in H^1(\mc{X};M)$ and $\varepsilon \geq n^{-1/d}$, then the bias term $B_{n,\varepsilon}^{(s)}(f)$ satisfies
+	\begin{equation}
+	\label{eqn:graph_seminorm_bias_2}
+	\mathbb{E}\bigl[|B_{n,\varepsilon}^{(s)}(f)|\bigr] \leq \frac{C\varepsilon^2}{\delta n\varepsilon^{2 + d}}M^2.
+	\end{equation}
+\end{lemma}
+Then~\ref{pf:graph_seminorm_ho_1} follows immediately from Lemma~\ref{lem:graph_seminorm_bias}, by Markov's inequality.
+\paragraph{Proof (of Lemma~\ref{lem:graph_seminorm_bias}).}
+We begin by introducing some notation.
+We will use bold notation $\bj = (j_1,\ldots,j_s)$ for a vector of indices where $j_i \in [n]$ for each $i$.
+We write $[n]^s$ for the collection of all such vectors, and $(n)^s$ for the subset of such vectors with no repeated indices.
+Finally, we write $D_if$ for a kernelized difference operator,
+\begin{equation*}
+D_if(x) := \bigl(f(x) - f(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr),
+\end{equation*}
+and we let $D_{\bj}f(x) := \bigl(D_{j_1}\circ \cdots \circ D_{j_s}f\bigr)(x)$.
+
+With this notation in hand, it is easy to represent $\dotp{L_{n,\varepsilon}^sf}{f}_{n}$ as the sum of a U-statistic and a bias term,
+\begin{align*}
+\dotp{L_{n,\varepsilon}^sf}{f}_{n} & = \frac{1}{n} \sum_{i = 1}^{n} L_{n,\varepsilon}^sf(X_i) \cdot f(X_i) \\
+& = \underbrace{\frac{1}{n^{s + 1}\varepsilon^{s(d + 2)}} \sum_{{i\bf j} \in (n)^{s + 1}}D_{\bj}f(X_i) \cdot f(X_i)}_{=:U_{n,\varepsilon}^{(s)}(f)} + \underbrace{\frac{1}{n^{s + 1}\varepsilon^{s(d + 2)}} \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot f(X_i)}_{=:B_{n,\varepsilon}^{(s)}(f)}
+\end{align*}
+When the indices of $i\bj$ are all distinct, it follows straightforwardly from the law of iterated expectation that
+\begin{equation*}
+\mathbb{E}[ D_{\bj}f(X_i) \cdot f(X_i)] = \varepsilon^{s(d + 2)}\mathbb{E}[L_{P,\varepsilon}^sf(X_i) \cdot f(X_i)] = \dotp{L_{P,\varepsilon}^sf}{f}_{P}, 
+\end{equation*}
+which in turn implies $\mathbb{E}[U_{n,\varepsilon}^{(s)}(f)] = (n - s - 1)!/n! \cdot \dotp{L_{P,\varepsilon}^sf}{f}_P$.
+
+It remains to show~\eqref{eqn:graph_seminorm_bias_2}. By adding and subtracting $f(X_{\bj_1})$, we obtain by symmetry that
+\begin{equation*}
+\sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot f(X_i) = \frac{1}{2} \cdot \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot \bigl(f(X_i) - f(X_{\bj_1})\bigr),
+\end{equation*}
+and consequently
+\begin{equation*}
+\Ebb\Bigl[\sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot f(X_i)\Bigr] \leq \frac{1}{2} \cdot \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} \Ebb\Bigl[\bigl|D_{\bj}f(X_i)\bigr| \cdot \bigl|f(X_i) - f(X_{\bj_1})\bigr|\Bigr].
+\end{equation*}
+In Lemma~\textcolor{red}{(?)}, we show that if $f \in H^1(\mc{X};M)$, then for any $i\bj \in [n]^{s + 1}$ which contains a total of $k + 1$ distinct indices, 
+\begin{equation*}
+\Ebb\Bigl[\bigl|D_{\bj}f(X_i)\bigr| \cdot \bigl|f(X_i) - f(X_{\bj_1})\bigr|\Bigr] \leq C_1 \varepsilon^{2 + kd} M^2.
+\end{equation*}
+This shows us that the expectation of $|B_{n,\varepsilon}^s(f)|$ can bounded from above by the sum over several different terms, as follows:
+\begin{align*}
+\Ebb\Bigl[|B_{n,\varepsilon}^s(f)|\Bigr] & \leq C_1\frac{\varepsilon^2}{n\varepsilon^{2s}}M^2 \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} \frac{1}{(n\varepsilon^d)^s}  \varepsilon^{(|i\bj| - 1)d} \\
+& \leq C_1\frac{\varepsilon^2}{n\varepsilon^{2s}}M^2  \sum_{k = 1}^{s - 1} \frac{(n\varepsilon^d)^k}{(n\varepsilon^d)^s}n.
+\end{align*}
+Finally, we note that by assumption $n\varepsilon^d \geq 1$, so that in the above sum the factor of $(n\varepsilon^d)^k$ is largest when $k = s- 1$. We conclude that
+\begin{equation*}
+\Ebb\Bigl[|B_{n,\varepsilon}^s(f)|\Bigr] \leq C_1 (s - 1) \frac{\varepsilon^2}{n\varepsilon^{2s + d}}M^2,
+\end{equation*}
+which is the desired result.
 
@@
 \end{equation}
-In the above, we have used that $|d_x^jf(z)| \leq |d^jf(x)|$ for all $z \in B(0,1)$, and the normalization $\int \eta(\|z\|) \,dz = 1$. 
+In the above, we have used both that $|d_x^jf(z)| \leq |d^jf(x)|$ for all $z \in B(0,1)$, and that the kernel is normalized so that $\int \eta(\|z\|) \,dz = 1$. 
 Combining this with~\eqref{pf:approximation_error_nonlocal_laplacian_4}, we conclude that
@@
 \int_{\mc{X}_{\varepsilon}} |G_2(x)|^2 \,dx & \leq C \Bigl(\varepsilon^{s - 2}\|p\|_{C^{s - 1}(\mc{X})}\Bigr)^2 \sum_{j = 1}^{s - 1} \int_{\mc{X}_{\varepsilon}}\biggl(\frac{1}{j!} \int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(z)\Bigr| \cdot \Bigl|\eta(\|z\|)\Bigr| \,dz\biggr)^2 \,dx \\
-& \leq C \Bigl(\varepsilon^{s - 2}\|p\|_{C^{s - 1}(\mc{X})} \sum_{j = 1}^{s - 1}\Bigr)^2 \|d^ju\|_{\Leb^2(\mc{X_{\varepsilon}})}^2,
+& \leq C \Bigl(\varepsilon^{s - 2}\|p\|_{C^{s - 1}(\mc{X})}\Bigr)^2 \sum_{j = 1}^{s - 1} \|d^ju\|_{\Leb^2(\mc{X_{\varepsilon}})}^2,
 \end{align*}
@@
 
+\textcolor{red}{(TODO): The language ``...for a Borel set $U \subseteq \mc{X}$...'' should be checked.}
+
 \begin{lemma}
@@
 	Assume Model~\ref{def:model_flat_euclidean}.
-Suppose $f \in L^2(U;M)$ for an open set $U \subseteq \mc{X}$, and let $L_{P,\varepsilon}$ be defined with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_flat_euclidean}.
-Then there exists a constant $C$ which does not depend on $f$ such that
+Suppose $f \in L^2(U;M)$ for a Borel set $U \subseteq \mc{X}$, and let $L_{P,\varepsilon}$ be defined with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_flat_euclidean}.
+Then there exists a constant $C$ which does not depend on $f$ or $M$ such that
 	\begin{equation}
@@
 \end{lemma}
+
+\begin{lemma}
+	\label{lem:graph_seminorm_bias2}
+	Assume Model~\ref{def:model_flat_euclidean}.
+Suppose $f \in H^1(\mc{X};M)$, and let $D_if$ be defined with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_flat_euclidean}.
+Then there exists a constant $C$ which does not depend on $f$ or $M$, such that for any $i \in [n]$ and $\bj \in [n]^s$,
+	\begin{equation*}
+	\Ebb\bigl[|D_{\bj}f(X_i)| \cdot |f(X_i) - f(X_{\bj_1})|\bigr] \leq C \varepsilon^{2 + dk}M^2,
+	\end{equation*}
+	where $k + 1$ is the number of distinct indices in $i\bj$. 
+\end{lemma}
+
 \paragraph{Proof (of Lemma~\ref{lem:l2estimate_nonlocal_laplacian}).}
@@
 & \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx.
-\label{pf:approximation_error_nonlocal_laplacian_2}
+\label{pf:l2estimate_nonlocal_laplacian_1}
 \end{align}
 Finally, using Fubini's Theorem we determine that
-\begin{equation*}
-\int_{U} \int{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx = 2 \int_{U} \int_{U} |f(x)|^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx \leq 2 \varepsilon^d \int_{U} |f(x)|^2 \,dx = 2\varepsilon^d \|f\|_{\Leb^2(U)}^2
-\end{equation*}
-and in combination with \eqref{pf:approximation_error_nonlocal_laplacian_2} we conclude that
+\begin{equation}
+\int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx = 2 \int_{U} \int_{U} |f(x)|^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx \leq 2 \varepsilon^d \int_{U} |f(x)|^2 \,dx = 2\varepsilon^d \|f\|_{\Leb^2(U)}^2,
+\label{pf:l2estimate_nonlocal_laplacian_2}
+\end{equation}
+and by combining~\eqref{pf:l2estimate_nonlocal_laplacian_1} and~\eqref{pf:l2estimate_nonlocal_laplacian_2} we conclude that
 \begin{equation*}
@@
 \end{equation*}
+
+\paragraph{Proof (of Lemma~\ref{lem:dirichlet_estimate_nonlocal_laplacian}).}
+We have
+\begin{equation*}
+E_{P,\varepsilon}(f) = \frac{1}{\varepsilon^{2 + d}} \int_{U} \int_{U} \bigl(f(x) - f(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) \leq \frac{2p_{\max}^2}{\varepsilon^{2 + d}} \int_{U} \int_{U} \bigl(|f(x)|^2 + |f(x')|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx,
+\end{equation*}
+and the claim follows from~\eqref{pf:l2estimate_nonlocal_laplacian_2}.
+
+\paragraph{Proof (of Lemma~\ref{lem:graph_seminorm_bias2}).}
+Let $G_{n,\varepsilon}[X_{i\bj}]$ be the subgraph induced by vertices $X_i, X_{\bj_1},\ldots,X_{\bj_s}$.
+We make two observations.
+First, in order for $|D_{\bj}f(X_i)| \cdot |f(X_i)  - f(X_j)|$ to be non-zero, it must be the case that the subgraph $G_{n,\varepsilon}[X_{i\bj}]$ is connected.
+Second, noting that for any indices $i$ and $j$,
+\begin{equation*}
+|D_{ij}f(x)| \leq \Bigl(|D_jf(X_i)| + |D_jf(x)|\Bigr)\|\eta\|_{\infty}, 
+\end{equation*}
+a straightforward inductive argument implies that 
+\begin{equation*}
+|D_{\bj}f(X_i)| \leq s\|\eta\|_{\infty}^s \sum_{j \in i\bj} |D_{\bj_s}f(X_j)|.
+\end{equation*}
+Combining these two observations, we reduce the task to upper bounding the product of two (first-order) differences,
+\begin{align*}
+\mathbb{E}\Bigl[ |D_{\bj}f(X_i)| |f(X_i) - f(X_{\bj_1})| \Bigr] & = \mathbb{E}\Bigl[ |D_{\bj}f(X_i)| |f(X_i) - f(X_{\bj_1})| \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr] \\
+& \leq s\|\eta\|_{\infty}^s \sum_{j \in i\bj} \mathbb{E}\Bigl[ |D_{\bj_s}f(X_j)| \cdot |f(X_i) - f(X_{\bj_1})| \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\}   \Bigr] \\
+& \leq s\|\eta\|_{\infty}^s \sum_{j \in i\bj} \mathbb{E}\Bigl[ |f(X_j) - f(X_{\bj_s})| \cdot |f(X_i) - f(X_{\bj_1})| \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\}   \Bigr]
+\end{align*}
+Next, from the Cauchy-Schwarz inequality we have that for any $j \in \bj$,
+\begin{align*}
+& \mathbb{E}\Bigl[ |f(X_j) - f(X_{\bj_s})| \cdot |f(X_i) - f(X_{\bj_1})| \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\}   \Bigr] \\
+& \quad \leq \sqrt{\mathbb{E}\Bigl[ |f(X_j) - f(X_{\bj_s})|^2 \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr]} \cdot \sqrt{\mathbb{E}\Bigl[ |f(X_j) - f(X_{\bj_s})|^2 \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr]} \\
+& \quad = \mathbb{E}\Bigl[ |f(X_j) - f(X_{i})|^2 \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr],
+\end{align*}
+with the equality following since each $X_i$ are identically distributed. Marginalizing out the contribution of all indices in $\bj$ not equal to $i$ or $j$ gives
+\begin{align}
+\mathbb{E}\Bigl[ |f(X_j) - f(X_{i})|^2 \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr] & \leq \bigl((s + 1)p_{\max}\nu_d\varepsilon^d\bigr)^{|i\bj\setminus\{j \cup i\}|} \cdot \mathbb{E}\Bigl[ |f(X_j) - f(X_{i})|^2 \1\{\|X_i - X_j\| \leq \varepsilon\}\Bigr] \nonumber \\
+& \leq \bigl((s + 1)p_{\max}\nu_d\varepsilon^d\bigr)^{|i\bj\setminus\{j \cup i\}|} \cdot p_{\max}^2 \nu_d \varepsilon^{2 + d} M^2 \label{pf:graph_seminorm_bias2_1}
+\end{align}
+with the second inequality following from the proof of Lemma~1 in \textcolor{blue}{(Green 2021).} Finally, we notice that $|i\bj\setminus \{i \cup j\}| + 1 = k$, so that~\eqref{pf:graph_seminorm_bias2_1} gives the desired result.
 

### 2021-04-04 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
Found one last typo.
few more changes mostly to Appendix + questions

@@
 where in the above the infimum is over all level-$\alpha$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Clearly, the minimax critical radius $\epsilon$ depends on $\alpha$ and $b$.
-However, we adopt the typical convention of treating $\alpha \in (0,1)$ and $b \geq 1$ as small but fixed positive constants; hence they will not affect the testing error rates, and we suppress them notationally.} See \citet{ingster82,ingster87,ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
+However, we adopt the typical convention of treating $\alpha \in (0,1)$ and $b \geq 1$ as small but fixed positive constants; hence they will not affect the testing error rates, and we suppress them notationally.}
+
+The classical approach to hypothesis testing typically focuses on designing test statistics, and studying their (limiting) distribution in order to ensure control of the Type I error.
+In many cases the Type II error (or risk in our terminology) is not emphasized, or the risk of the test against fixed or directional alternatives (i.e.
+alternatives which deviate from the null in a fixed direction) is studied.
+In contrast, in the minimax paradigm the (uniform or worst-case) risk against a large collection of alternatives is the central focus.
+See \citet{ingster82,ingster87,ingster2012,ariascastro2018,balakrishnan2019,balakrishnan2018hypothesis} for a more extended treatment of the minimax paradigm in nonparametric testing, and for a discussion of its advantages (and disadvantages) over other approaches to studying hypothesis tests.
 
@@
 \begin{equation*}
-\wh{\varphi} = \1\bigl\{\wh{T} \leq \wh{t}_{\alpha}\bigr\}.
+\wh{\varphi} = \1\bigl\{\wh{T} > \wh{t}_{\alpha}\bigr\}.
 \end{equation*} 

### 2021-04-04 manuscripts/laplacian_smoothing/arXiv/appendix.tex
few more changes mostly to Appendix + questions

@@
 	\label{eqn:optimal_transport_2}
-	\|\wt{p}_n - p\|_{\infty} \leq A_0\bigl(\wt{\delta} + \theta\bigr)
+	\|\wt{p}_n - p\|_{\infty} \leq A_0\bigl(\wt{\delta} + \theta\bigr).
 	\end{equation}
@@
 where the last inequality follows by taking $C_6 \geq C_7 \sqrt{p_{\max}/p_{\min}}$.
+\sbcomment{I'm confused about the last inequality.
+Are you missing squares on the last term, i.e.
+$(\Ebb\bigl[\norm{f}_n^2\bigr])^2$ ?
+I think the problem repeats below but I could be just missing something.}
 
@@
 
-\paragraph{Bounds on $\Leb^2(\Xset)$ error under Lipschitz assumption.}
+\paragraph{Bounds on $\Leb^2(\Xset)$ error under Lipschitz assumption.} \sbcomment{This is a nice result but where do you reference it or use it?}
 Let $V_1,\ldots,V_n$ denote the Voronoi tesselation of $\Xset$ with respect to $X_1,\ldots,X_n$. Extend $\wh{f}$ over $\Xset$ by taking it piecewise constant over the Voronoi cells, i.e.

### 2021-04-05 manuscripts/laplacian_eigenmaps/appendix.tex
Added conditional-on-design estimation and testing error statements.

@@
 In this section, we adopt the fixed design perspective; or equivalently, condition on $X_i = x_i$ for $i = 1,\ldots,n$.
-Let $G = \bigl([n],W\bigr)$ be a fixed graph on $\{1,\ldots,n\}$ with Laplacian matrix $L = \sum_{k = 1}^{n}\lambda_k v_k v_k^{\top}$.
+Let $G = \bigl([n],W\bigr)$ be a fixed graph on $\{1,\ldots,n\}$ with Laplacian matrix $L = \frac{1}{n}\sum_{k = 1}^{n}\lambda_k v_k v_k^{\top}$; the eigenvectors have unit empirical norm, $\|v_k\|_n^2 = 1$.
 The randomness thus all comes from the responses
@@
 
+\subsection{Upper bound on Estimation Error of Laplacian Eigenmaps}
+
+\begin{lemma}
+	\label{lem:fixed_graph_estimation}
+	For any $s \in \mathbb{N} \setminus \{0\}$, the Laplacian eigenmaps estimator $\wh{f}$ of~\eqref{eqn:laplacian_eigenmaps_estimator} satisfies
+	\begin{equation*}
+	\|\wh{f} - f_0\|_n^2 \leq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} + \frac{5K}{n},
+	\end{equation*}
+	with probability at least $1 - \exp(-K)$. 
+\end{lemma}
+\begin{proof}
+	By the triangle inequality,
+	\begin{equation}
+	\label{pf:fixed_graph_estimation_1}
+	\|\wh{f} - f_0\|_n^2 \leq 2\Bigl(\|\mathbb{E}\wh{f} - f_0\|_n^2 + \|\wh{f} - \mathbb{E}\wh{f}\|_n^2\Bigr).
+	\end{equation}
+	The first term in~\eqref{pf:fixed_graph_estimation_1} (bias) is non-random, since the design is fixed. The expectation $\mathbb{E}\wh{f} = \sum_{k = 1}^{K} \dotp{v_k}{f_0}_n v_k$, so that
+	\begin{equation*}
+	\|\mathbb{E}\wh{f} - f_0\|_n^2 = \Bigl\|\sum_{k = K + 1}^{n} \dotp{v_k}{f_0}_n v_k\Bigr\|_n^2 = \sum_{k = K + 1}^n \dotp{v_k}{f_0}_n^2.
+	\end{equation*}
+	In the above, the last equality relies on the fact that $v_k$ are orthonormal. Using the fact that the eigenvalues are in increasing order, we obtain that
+	\begin{equation*}
+	\sum_{k = K + 1}^n \dotp{v_k}{f_0}_n^2 \leq \frac{1}{\lambda_K^s} \sum_{k = K + 1}^n \lambda_k^s \dotp{v_k}{f_0}_n^2 \leq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s}.
+	\end{equation*}
+
+	The second term in~\eqref{pf:fixed_graph_estimation_1} (variance) is random. Again using the orthonormality of the eigenvectors $v_k$, we have
+	\begin{equation*}
+	\|\wh{f} - \mathbb{E}\wh{f}\|_n^2 = \sum_{k = 1}^{K} \dotp{v_k}{\varepsilon}_n^2.
+	\end{equation*}
+	Thus $\|\wh{f} - \mathbb{E}\wh{f}\|_n^2$ is equal to $1/n$ times a $\chi^2$ distribution with $K$ degrees of freedom. Consequently, it follows from a result of \citep{laurent00} that
+	\begin{equation*}
+	\Pbb\Bigl(\|\wh{f} - \mathbb{E}\wh{f}\|_n^2 \geq \frac{K}{n} + 2\frac{\sqrt{K}}{n}\sqrt{t} + \frac{2t}{n}\Bigr) \leq \exp(-t).
+	\end{equation*}
+	Setting $t = K$ completes the proof of the Lemma.
+\end{proof}
+
+\subsection{Upper bound on Testing Error of Laplacian Eigenmaps}
+
 Let $\wh{T} = \sum_{k = 1}^{K} \dotp{{\bf Y}}{v_k}_n^2$, and let $\varphi = \1\{\wh{T} \geq t_a\}$. In the following Lemma, we upper bound the Type I and Type II error of the test $\varphi$.
@@
 		\begin{equation*}
-		\sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 \geq 16\frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr].
+		\|f_0\|_n^2 \geq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} + 16\frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr],
 		\end{equation*}
-		Then $\Ebb_{f_0}[1 - \phi] \leq b$.
+		for some $s \in \mathbb{N}\setminus \{0\}$. Then $\Ebb_{f_0}[1 - \phi] \leq b$.
 	\end{itemize}

### 2021-04-05 manuscripts/laplacian_smoothing/arXiv/appendix.tex
Responded to Siva's comments.

--- 2021-03-09 manuscripts/laplacian_smoothing/arXiv/appendix.tex
@@
 \end{equation*}
-We will use the Sobolev inequalities as a tool to show that $\|f\|_{\Leb^4(\Xset)}/n \leq \bigl(\mathbb{E}[\|f\|_n^2]\bigr)^2/(b^2p_{\max})$, whence the claim of the Lemma is shown.
+We will use the Sobolev inequalities as a tool to show that $\|f\|_{\Leb^4(\Xset)}^4/n \leq \bigl(\mathbb{E}[\|f\|_n^2]\bigr)^2/(b^2p_{\max})$, whence the claim of the Lemma is shown.
 The nature of the inequalities we use depend on the value of $d$.
@@
 \begin{equation*}
-p_{\max} \frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq C_7^2 M^2p_{\max} \cdot \frac{\norm{f}_{\Leb^2(\Xset)}^2}{n} \leq \frac{C_7p_{\max}}{C_6^2 b^2} \norm{f}_{\Leb^2(\Xset)}^4 \leq \frac{\Ebb\bigl[\norm{f}_n^2\bigr]}{b^2},
+p_{\max} \frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq C_7^2 M^2p_{\max} \cdot \frac{\norm{f}_{\Leb^2(\Xset)}^2}{n} \leq \frac{C_7p_{\max}}{C_6^2 b^2} \norm{f}_{\Leb^2(\Xset)}^4 \leq \frac{\bigl(\Ebb\bigl[\norm{f}_n^2\bigr]\bigr)^2}{b^2},
 \end{equation*}
 where the last inequality follows by taking $C_6 \geq C_7 \sqrt{p_{\max}/p_{\min}}$. 
+%\sbcomment{I'm confused about the last inequality.
+Are you missing squares on the last term, i.e.
+$(\Ebb\bigl[\norm{f}_n^2\bigr])^2$ ?
+I think the problem repeats below but I could be just missing something.}
+% AG: Siva was correct, I was missing a square on the last term, here and below. Resolved, here and below.
 
@@
 \begin{equation*}
-p_{\max} \frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq \norm{f}_{\Leb^2(\Xset)}^4p_{\max} \cdot \left(\frac{C_7\norm{f}_{H^1(\Xset)}}{n^{1/d}\norm{f}_{\Leb^2(\Xset)}}\right)^{d} \leq \frac{C_7^dp_{\max}\norm{f}_{\Leb^2(\Xset)}^4}{C_6^db^{d}} \leq \frac{\Ebb\bigl[\norm{f}_n^2\bigr]}{b^2}.
+p_{\max} \frac{\norm{f}_{\Leb^4(\Xset)}^4}{n} \leq \norm{f}_{\Leb^2(\Xset)}^4p_{\max} \cdot \left(\frac{C_7\norm{f}_{H^1(\Xset)}}{n^{1/d}\norm{f}_{\Leb^2(\Xset)}}\right)^{d} \leq \frac{C_7^dp_{\max}\norm{f}_{\Leb^2(\Xset)}^4}{C_6^db^{d}} \leq \frac{\bigl(\Ebb\bigl[\norm{f}_n^2\bigr]\bigr)^2}{b^2}.
 \end{equation*}
@@
 \end{equation} 
-Putting $q = 2/(1 - t)$, we have that $\|f\|_{\Leb^{q}(\mc{X})} \leq C_7 \cdot M$, and it follows from derivations similar to those in Case 2 that $\|f\|_{\Leb^4(\Xset)}/n \leq \bigl(\mathbb{E}[\|f\|_n^2]\bigr)^2/(b^2p_{\max})$ when $C_6 \geq C_7 \sqrt{p_{\max}/p_{\min}}$.
+Putting $q = 2/(1 - t)$, we have that $\|f\|_{\Leb^{q}(\mc{X})} \leq C_7 \cdot M$, and it follows from derivations similar to those in Case 2 that $\|f\|_{\Leb^4(\Xset)}^4/n \leq \bigl(\mathbb{E}[\|f\|_n^2]\bigr)^2/(b^2p_{\max})$ when $C_6 \geq C_7 \sqrt{p_{\max}/p_{\min}}$.
 
@@
 \begin{equation*}
-\frac{1}{n}\|f\|_{\Leb^4(\Xset)} = \frac{1}{n}\lim_{k \to \infty} \|f_k\|_{\Leb^4(\Xset)} \leq \frac{1}{b^2p_{\max}} \lim_{k \to \infty} \bigl(\mathbb{E}[\|f_k\|_n^2]\bigr)^2 =  \frac{1}{b^2p_{\max}}\bigl(\mathbb{E}[\|f\|_n^2]\bigr)^2.
+\frac{1}{n}\|f\|_{\Leb^4(\Xset)}^4 = \frac{1}{n}\lim_{k \to \infty} \|f_k\|_{\Leb^4(\Xset)}^4 \leq \frac{1}{b^2p_{\max}} \lim_{k \to \infty} \bigl(\mathbb{E}[\|f_k\|_n^2]\bigr)^2 =  \frac{1}{b^2p_{\max}}\bigl(\mathbb{E}[\|f\|_n^2]\bigr)^2.
 \end{equation*}
@@
 \paragraph{Bounds on $\Leb^2(\Xset)$ error under Lipschitz assumption.} 
+% \sbcomment{This is a nice result but where do you reference it or use it?}
+% AG: Resolved.
 Let $V_1,\ldots,V_n$ denote the Voronoi tesselation of $\Xset$ with respect to $X_1,\ldots,X_n$. Extend $\wh{f}$ over $\Xset$ by taking it piecewise constant over the Voronoi cells, i.e.
@@
 Therefore Proposition~\ref{prop:out_of_sample_error} together with Theorem~\ref{thm:laplacian_smoothing_estimation1} implies that with high probability, $\wh{f}$ achieves the nearly-optimal (up to a factor of $\log n$) estimation rates out-of-sample error---that is, $\|\wh{f} - f_0\|_{\Leb^2(\Xset)}^2 \leq C \log(n) M^{2d/(2 + d)}n^{-2/(2+d)}$---as long as $M \leq Cn^{1/d}$.
+This justifies one of our remarks after Theorem~\ref{thm:laplacian_smoothing_estimation1}.
 

### 2021-04-05 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
Responded to Siva's comments.

@@
 Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius of the Sobolev ball is also optimal.
-Rates for thin-plate splines with $d \geq 2$ assume the estimator $\wt{f}$ interpolates the responses, $\wt{f}(X_i) = Y_i$ for $i = 1,\ldots,n$; see the discussion in Section~\ref{sec:summary}.
-We use ``$1$'' to indicate inconsistency (error not converging to 0).
-Furthermore, when $\Xset$ is an $m$-dimensional manifold embedded in $\Rd$, all Laplacian smoothing results hold with $d$ replaced by $m$, without any change to the method itself.}
+The rates for thin-plate splines with $d \geq 2$ assume the estimator \smash{$\wt{f}$} interpolates the responses, \smash{$\wt{f}(X_i) = Y_i$} for $i = 1,\ldots,n$; see the discussion in Section~\ref{sec:summary}.
+Here, we use ``$1$'' to indicate inconsistency (error not converging to 0).
+Lastly, when $\Xset$ is an $m$-dimensional manifold embedded in $\Rd$, all Laplacian smoothing results hold with $d$ replaced by $m$, without any change to the method itself.}
 	\label{tbl:estimation_rates}
@@
 	\caption{Summary of testing rates over first-order Sobolev balls; black, red, and bold fonts are used as in Table~\ref{tbl:estimation_rates}.
-Rates for thin-plate splines with $d \geq 2$ assume the test statistic $\wt{T}$ is computed with respect to an estimator $\wt{f}$ that interpolates the responses, $\wt{f}(X_i) = Y_i$ for $i = 1,\ldots,n$.
+The rates for thin-plate splines with $d \geq 2$ assume the test statistic \smash{$\wt{T}$} is computed using an \smash{$\wt{f}$} that interpolates the responses, $\wt{f}(X_i) = Y_i$ for $i = 1,\ldots,n$.
 Rates for $d \geq 4$ assume that $f_0 \in \Leb^4(\Xset,M)$.
-When $\Xset$ is an $m$-dimensional manifold embedded in $\Rd$, all rates hold with $d$ replaced by $m$.}
+Lastly, when $\Xset$ is an $m$-dimensional manifold embedded in $\Rd$, all rates hold with $d$ replaced by $m$.}
 	\label{tbl:testing_rates}
@@
 When \smash{$\wh{f}$} and $f_0$ are suitably smooth, tools from empirical process theory (see e.g., Chapter 14 of \citet{wainwright2019}) or approximation theory (e.g., Section 15.5 of \citet{johnstone2011}) guarantee that the $\Leb^2(\Xset)$ error is not too much greater than its in-sample counterpart.
-In fact, as shown in Appendix~\ref{sec:graph_sobolev_seminorm}, if $f_0$ is Lipschitz smooth and we extend \smash{$\wh{f}$} to be piecewise constant over the Voronoi tessellation induced by $X_1,\ldots,X_n$, then the out-of-sample error \smash{$\|\wh{f}-f_0\|_{\Leb^2(\Xset)}$} is within a negligible factor of the in-sample error \smash{$\|\wh{f}-f_0\|_n$}.
+In fact, as shown in Appendix~\ref{subsec:laplacian_smoothing_estimation1_pf}, if $f_0$ is Lipschitz smooth and we extend \smash{$\wh{f}$} to be piecewise constant over the Voronoi tessellation induced by $X_1,\ldots,X_n$, then the out-of-sample error \smash{$\|\wh{f}-f_0\|_{\Leb^2(\Xset)}$} is within a negligible factor of the in-sample error \smash{$\|\wh{f}-f_0\|_n$}.
 We leave analysis of the Sobolev case to future work.
@@
 %(Clearly \smash{$\wh{\varphi}$} depends on $b$, but we suppress this notationally.) 
-We show in Appendix~\ref{sec:fixed_graph_error_bounds} that $\wh{f}$ is a level-$\alpha$ test.
-In the next theorem, we upper bound the worst-case risk $R_n(\wh{\varphi},H^1(\mc{X},M),\epsilon)$ of $\wh{\varphi}$, whenever $\epsilon$ is at least (a constant times) the critical radius given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
+We show in Appendix~\ref{sec:fixed_graph_error_bounds} that \smash{$\wh{f}$} is a level-$\alpha$ test.
+In the next theorem, we upper bound the worst-case risk \smash{$R_n(\wh{\varphi},H^1(\mc{X},M),\epsilon)$} of \smash{$\wh{\varphi}$}, whenever $\epsilon$ is at least (a constant times) the critical radius given in~\eqref{eqn:sobolev_space_testing_critical_radius}.
 For this to hold, we will require a tighter range of scalings for the graph radius $r$.
@@
 	\label{thm:laplacian_smoothing_testing}
-	Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ with $d < 4$, and where $M \leq M_{\max}(d)$.
+	Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ with $d < 4$, and $M \leq M_{\max}(d)$.
 Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume \smash{$G_{n,r}$} is computed with a kernel $K$ satisfying \ref{asmp:kernel}.
-There are constants $N,C,C_1,c_1>0$ such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_testing}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$}, satisfies the following: for any $b \geq 1$, if
+There exist constants $N,C,C_1,c_1>0$ such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_testing}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$}, satisfies the following: for any $b \geq 1$, if
 	\begin{equation}
 	\label{eqn:laplacian_smoothing_testing}
-	\epsilon^2 \geq C M^{2d/(4 + d)} n^{-4/(4 + d)}\Bigl(b^2 + b\sqrt{\frac{1}{\alpha}}\Bigr) ,
+	\epsilon^2 \geq C M^{2d/(4 + d)} n^{-4/(4 + d)}\biggl(b^2 + b\sqrt{\frac{1}{\alpha}}\biggr) ,
 	\end{equation} 
-	then the worst-case risk is upper bounded, $R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + C_1n\exp(-c_1nr^d)\bigr)$.
+	then the worst-case risk satisfies the upper bound: \smash{$R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + C_1n\exp(-c_1nr^d)\bigr)$}.
 \end{theorem}
@@
 \end{equation*}
-and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies e.g., Dirichlet boundary conditions.} To be clear, this argument does not formally imply that the Laplacian eigenmaps estimator $\wh{f}$ and the thin-plate spline estimator $\wt{f}$ are close (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of continuous functions with weak derivatives in $\Leb^2(\Xset)$).
+and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies Neumann boundary conditions.} To be clear, this argument does not formally imply that the Laplacian eigenmaps estimator \smash{$\wh{f}$} and the thin-plate spline estimator \smash{$\wt{f}$} are close (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of continuous functions with weak derivatives in $\Leb^2(\Xset)$).
 But it does seem to suggest that the two estimators should behave somewhat similarly.

### 2021-04-27 manuscripts/laplacian_eigenmaps/appendix.tex
Massive update to Laplacian Eigenmaps draft. Too many changes to list, but all generally responding to Ryan's comments.

@@
 	For any $s \in \mathbb{N} \setminus \{0\}$, the Laplacian eigenmaps estimator $\wh{f}$ of~\eqref{eqn:laplacian_eigenmaps_estimator} satisfies
-	\begin{equation*}
+	\begin{equation}
+	\label{eqn:fixed_graph_estimation}
 	\|\wh{f} - f_0\|_n^2 \leq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} + \frac{5K}{n},
-	\end{equation*}
+	\end{equation}
 	with probability at least $1 - \exp(-K)$. 
 \end{lemma}
-\begin{proof}
+\paragraph{Proof (of Lemma~\ref{lem:fixed_graph_estimation}).}
 	By the triangle inequality,
@@
 	\end{equation}
-	The first term in~\eqref{pf:fixed_graph_estimation_1} (bias) is non-random, since the design is fixed. The expectation $\mathbb{E}\wh{f} = \sum_{k = 1}^{K} \dotp{v_k}{f_0}_n v_k$, so that
+	The first term in~\eqref{pf:fixed_graph_estimation_1} (approximation error) is non-random, since the design is fixed.
+The expectation $\mathbb{E}\wh{f} = \sum_{k = 1}^{K} \dotp{v_k}{f_0}_n v_k$, so that
 	\begin{equation*}
@@
 
-	The second term in~\eqref{pf:fixed_graph_estimation_1} (variance) is random. Again using the orthonormality of the eigenvectors $v_k$, we have
-	\begin{equation*}
-	\|\wh{f} - \mathbb{E}\wh{f}\|_n^2 = \sum_{k = 1}^{K} \dotp{v_k}{\varepsilon}_n^2.
+	The second term in~\eqref{pf:fixed_graph_estimation_1} (estimation error) is random.
+Observe that $\dotp{v_k}{\varepsilon}_n \overset{d}{=} Z_k/\sqrt{n}$, where $(Z_1,\ldots,Z_n) \sim N(0,I_{n \times n})$.
+Again using the orthonormality of the eigenvectors $v_k$, we have
+	\begin{equation*}
+	\|\wh{f} - \mathbb{E}\wh{f}\|_n^2 = \sum_{k = 1}^{K} \dotp{v_k}{\varepsilon}_n^2 \overset{d}{=} \frac{1}{n}\sum_{k = 1}^{K} Z_k^2.
 	\end{equation*}
@@
 	\begin{equation*}
-	\Pbb\Bigl(\|\wh{f} - \mathbb{E}\wh{f}\|_n^2 \geq \frac{K}{n} + 2\frac{\sqrt{K}}{n}\sqrt{t} + \frac{2t}{n}\Bigr) \leq \exp(-t).
+	\Pbb\biggl(\|\wh{f} - \mathbb{E}\wh{f}\|_n^2 \geq \frac{K}{n} + 2\frac{\sqrt{K}}{n}\sqrt{t} + \frac{2t}{n}\biggr) \leq \exp(-t).
 	\end{equation*}
 	Setting $t = K$ completes the proof of the Lemma.
-\end{proof}
 
@@
 \begin{lemma}
+	\label{lem:fixed_graph_testing}
 	Suppose we observe $(Y_1,x_1),\ldots,(Y_n,x_n)$ according to~\eqref{eqn:fixed_graph_regression_model}.
@@
 		\item Suppose $f_0 \neq 0$ satisfies
-		\begin{equation*}
-		\|f_0\|_n^2 \geq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} + 16\frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr],
-		\end{equation*}
+		\begin{equation}
+		\label{eqn:fixed_graph_testing_critical_radius}
+		\|f_0\|_n^2 \geq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} + \frac{\sqrt{2K}}{n}\biggl[2\sqrt{\frac{1}{a}} + \sqrt{\frac{2}{b}} + \frac{32}{bn}\biggr],
+		\end{equation}
 		for some $s \in \mathbb{N}\setminus \{0\}$. Then $\Ebb_{f_0}[1 - \phi] \leq b$.
@@
 \end{lemma}
+\paragraph{Proof (of Lemma~\ref{lem:fixed_graph_testing}).}
+We first compute the expectation and variance of $\wh{T}$, then apply Chebyshev's inequality to upper bound the Type I and Type II error.
+
+\underline{\emph{Expectation}.}
+Recall that $\wh{T} = \sum_{k = 1}^{K} \dotp{Y}{v_k}_n^2$. Expanding the square gives
+\begin{equation*}
+\Ebb[\wh{T}] = \sum_{k = 1}^{K} \Ebb[\dotp{Y}{v_k}_n^2] = \sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 + \Ebb[2\dotp{f_0}{v_k}_n\dotp{\varepsilon}{v_k}_n + \dotp{\varepsilon}{v_k}_n^2] = \frac{K}{n} + \sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2.
+\end{equation*}
+Thus $\Ebb[\wh{T}] - t_a = \sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 - \sqrt{2K}/n \cdot \sqrt{1/a}$. Furthermore, it is a consequence of~\eqref{eqn:fixed_graph_testing_critical_radius} that 
+\begin{equation}
+\label{pf:fixed_graph_testing_1}
+\sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 - \frac{\sqrt{2K}}{n}\sqrt{1/a} \geq \|f_0\|_n^2 - \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} - \frac{\sqrt{2K}}{n}\sqrt{1/a} \geq \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \sqrt{\frac{2}{b}} + \frac{32}{bn}\biggr].
+\end{equation}
+
+\underline{\emph{Variance}.}
+Recall from the proof of Lemma~\ref{lem:fixed_graph_estimation} that $\dotp{\varepsilon}{v_k}_n \overset{d}{=} Z_k/\sqrt{n}$ for $(Z_1,\ldots,Z_n) \sim N(0,I_{n \times n})$.
+Expanding the square, and recalling that $\Cov[Z,Z^2] = 0$ for Gaussian random variables, we have that
+\begin{equation*}
+\Var\bigl[\dotp{{\bf Y}}{v_k}_n^2\bigr] = \Var\biggl[\frac{2}{n}\dotp{f_0}{v_k}_nZ_k + \frac{2}{n^2}Z_k^2\biggr] = \frac{4\dotp{f_0}{v_k}_n^2}{n} + \frac{2}{n^2}.
+\end{equation*}
+Moreover, since $\Cov[Z_k^2,Z_{\ell}^2] = 0$ for each $k = 1,\ldots,K$, we see that
+\begin{equation*}
+\Var\bigl[\wh{T}\bigr] = \sum_{k = 1}^{K} \Var\bigl[\dotp{{\bf Y}}{v_k}_n^2\bigr] = \frac{2K}{n^2} + \sum_{k = 1}^{K}\frac{4\dotp{f_0}{v_k}_n^2}{n}.
+\end{equation*}
+
+\underline{\emph{Bounds on Type I and Type II error}.}
+The upper bound on Type I error follows immediately from Chebyshev's inequality.
+
+The upper bound on Type II error also follows from Chebyshev's inequality.
+We observe that~\eqref{eqn:fixed_graph_testing_critical_radius} implies $\Ebb_{f_0}[\wh{T}] = t_a$, and apply Chebyshev's inequality to deduce
+\begin{equation*}
+\Pbb_{f_0}\bigl(\wh{T} < t_a\bigr) \leq \Pbb_{f_0}\Bigl(|\wh{T} - \Ebb_{f_0}[\wh{T}]|^2 > |\Ebb_{f_0}[\wh{T}] - t_a|^2\Bigr) \leq \frac{\Var\bigl[\wh{T}\bigr]}{\bigl[\Ebb_{f_0}[\wh{T}] - t_a\bigr]^2} = \frac{2K/n^2 + 4/n\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2}{\bigl[\Ebb_{f_0}[\wh{T}] - t_a\bigr]^2}.
+\end{equation*}
+Thus we have upper bounded the Type II error by the sum of two terms, each of which are no more than $1/(2b)$, as we now show.
+For the first term, after noting that~\eqref{pf:fixed_graph_testing_1} implies $\Ebb_{f_0}[\wh{T}] - t_a \geq \sqrt{2K}/n \cdot \sqrt{2/b}$, the upper bound follows:
+\begin{equation*}
+\frac{2K/n^2}{\bigl[\Ebb_{f_0}[\wh{T}] - t_a\bigr]^2} \leq \frac{b}{2}.
+\end{equation*}
+On the other hand, for the second term we use~\eqref{pf:fixed_graph_testing_1} in two ways: first to conclude that $\Ebb_{f_0}[\wh{T}] - t_a \geq 1/2 \cdot \sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2$, and second to obtain
+\begin{equation*}
+\frac{4\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2}{n\bigl[\Ebb_{f_0}[\wh{T}] - t_a\bigr]^2} \leq \frac{4\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2}{n\bigl(\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2/2\bigr)^2} \leq \frac{16}{n\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2} \leq \frac{b}{2}.
+\end{equation*}
+
+\section{Analysis of Spectral Series Estimator}
+\label{sec:analysis_spectral_series_estimator}
+In this section we prove Proposition~\ref{prop:spectral_series_estimation}.
+As mentioned in our main text, this proof relies on three facts: (i) a continuous embedding of $H_0^s(\mc{X})$ into $\mc{H}^s(\mc{X})$, (ii) the Weyl's Law asymptotic scaling $\lambda_k(\Delta_P) \asymp k^{2/d}$, and (iii) a local Weyl's Law scaling $\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2 \lesssim K$.
+(Recall that $(\lambda_k(\Delta_P),\psi_k)$ are defined to be solutions to~\eqref{eqn:laplacian_eigenfunction_neumann}).
+We record the precise results we need in the following three lemmas.
+
+The first result follows immediately from Lemma~17 of \textcolor{blue}{(Dunlop)}.
+Recall the spectral Sobolev class $\mc{H}^s(\mc{X})$ defined in~\eqref{eqn:spectral_sobolev_space}.
+Equip $\mc{H}^s(\mc{X})$ with the norm
+\begin{equation*}
+\|f\|_{\mc{H}^s(\mc{X})}^2 := \sum_{k = 1}^{\infty} \bigl[\dotp{f}{\psi_k}_P\bigr]^2 \cdot [\lambda_k(\Delta_P)]^2.
+\end{equation*}
+\begin{lemma}[cf. Lemma~17 of \textcolor{blue}{(Dunlop)}]
+	\label{lem:zero_trace_spectral_sobolev}
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $p \in C^{\infty}(\mc{X})$ and $\partial \mc{X} \in C^{1,1}$.
+Then for any $s \in \mathbb{N}$, we have $H_0^s(\mc{X}) \subseteq \mc{H}^s(\mc{X})$.
+Moreover, there exists a constant $C$ such that for any $f \in H_0^s(\mc{X})$ we have
+	\begin{equation}
+	\label{eqn:zero_trace_spectral_sobolev}
+	\|f\|_{\mc{H}^s(\mc{X})} \leq C\|f\|_{H^s(\mc{X})}.
+	\end{equation}
+\end{lemma}
+
+The particular version of Weyl's Law we use is also due to \textcolor{blue}{(Dunlop)}. 
+\begin{lemma}[cf. Lemma 28 of \textcolor{blue}{(Dunlop)}]
+	\label{lem:weyl}
+	Suppose Model~\ref{def:model_flat_euclidean}. Then there exist positive constants $c$ and $C$ such that for all $k \geq 2$, 
+	\begin{equation}
+	\label{eqn:weyl}
+	ck^{2/d} \leq \lambda_k(\Delta_P) \leq Ck^{2/d}.
+	\end{equation}
+\end{lemma}
+As we discuss in our main text, the asymptotic scaling $\lambda_k(\Delta_P) \asymp k^{2/d}$ also plays a key role in our analysis of Laplacian eigenmaps.
+We point out that although Lemma 28 of \textcolor{blue}{(Dunlop)} assumes $p \in C^{\infty}(\mc{X})$ and $\partial \mc{X} \in C^{1,1}$, the proof uses only the hypotheses of Model~\ref{def:model_flat_euclidean}: that $p$ is bounded away from $0$ and $\infty$, and that $\partial \mc{X}$ is Lipschitz.
+
+Finally, we need a local version of Weyl's Law, due to \textcolor{blue}{(H\"{o}rmander)}.
+\begin{lemma}[cf. Theorem 17.5.3 of \textcolor{blue}{H\"{o}rmander}]
+	\label{lem:local_weyl}
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $p \in C^{\infty}(\mc{X})$ and $\partial \mc{X} \in C^{1,1}$.
+Then there exists a positive constant $C$ such that for all $K \in \mathbb{N} \setminus {0}$,
+	\begin{equation}
+	\label{eqn:local_weyl}
+	\sup_{x \in \mc{X}}\biggl\{\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2\biggr\} \leq CK.
+	\end{equation}
+\end{lemma}
+Translated into our notation, Theorem 17.5.3 of \textcolor{blue}{(H\"{o}rmander)} says that
+\begin{equation*}
+\sup_{x \in \mc{X}}\biggl\{\sum_{k: \lambda_k(\Delta_P) \leq \lambda_K(\Delta_P)}  \bigl(\psi_k(x)\bigr)^2\biggr\} \leq C \bigl[\lambda_K(\Delta_P)\bigr]^{d/2},
+\end{equation*}
+and from here~\eqref{eqn:local_weyl} follows from~\eqref{eqn:weyl}.
+\textcolor{blue}{(H\"{o}rmander)} proves the result only when $\partial \mc{X} \in C^{\infty}$, but the proof goes through unchanged when $\partial \mc{X} \in C^1$, since $\partial \mc{X} \in C^1$ is sufficient to apply the relevant Sobolev embedding theorem (Lemma 17.5.2 of~\textcolor{blue}{(H\"{o}rmander)}) that is the key to proving Lemma~\ref{lem:local_weyl}.
+
+We note that the assumptions $p \in C^{\infty}(\mc{X})$ and $\partial\mc{X} \in C^{1,1}$ used in Lemmas~\ref{lem:zero_trace_spectral_sobolev} and~\ref{lem:local_weyl} can likely be removed.
+This would allow us to remove the same assumptions in the statement of Proposition~\ref{prop:spectral_series_estimation}.
+Since this does not change the main point of the proposition, we do not pursue the details further.
+
+\paragraph{Proof (of Proposition~\ref{prop:spectral_series_estimation}).}
+We decompose risk into squared bias and variance,
+\begin{equation}
+\label{pf:spectral_series_estimation_0}
+\Ebb \|\wt{f} - f_0\|_P^2 = \Ebb\| \Ebb[\wt{f}]  - f_0\|_P^2 + \Ebb\| \wt{f} - \Ebb[\wt{f}]\|_P^2.
+\end{equation}
+Then some standard arguments (which we give below) give the following upper bound:
+\begin{equation}
+\label{pf:spectral_series_estimation_1}
+\Ebb \|\wt{f} - f_0\|_P^2 \leq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\bigl[\lambda_{K + 1}(\Delta_P)\bigr]^s} + \frac{K}{n} + \frac{1}{n}\Ebb\Bigl[(f_0(X))^2 \cdot \sum_{k = 1}^{K} (\psi_k(X))^2\Bigr].
+\end{equation}
+The claim of the proposition follows from~\eqref{pf:spectral_series_estimation_1} and Lemmas~\ref{lem:zero_trace_spectral_sobolev}-\ref{lem:local_weyl}, upon taking $K = \floor{(M^2n)^{d/(2s + d)}}$.
+(Here we use the convention $\sum_{k = 1}^{0}c_k = 0$ for any sequence $(c_k)$.) \qed
+
+The upper bound~\eqref{pf:spectral_series_estimation_1} is to be compared with~\eqref{eqn:estimation_biasvariance}, which gives upper bounds on the (design-dependent) squared bias and variance of Laplacian eigenmaps.
+We see~\eqref{eqn:estimation_biasvariance} has terms analogous to the first two terms on the right hand side of~\eqref{pf:spectral_series_estimation_1}, but not the third.
+The relevance of this third term to $L^2$ estimation problems is discussed by~\cite{birge2008}.
+
+\paragraph{Proof (of \eqref{pf:spectral_series_estimation_1}).}
+Since $\{\psi_k\}$ are an orthonormal basis of $L^2(P)$, and $f_0 \in H_0^{s}(\mc{X}) \subseteq L^2(P)$, we have $f_0 = \sum_{k = 1}^{\infty} \dotp{f_0}{\psi_k}_P \psi_k$ in $L^2(P)$.
+This allows us to expand the squared bias in terms of squared Fourier coefficients of $f_0$, and gives the following upper bound,
+\begin{equation*}
+\|f_0 - \Ebb{\wt{f}}\|_P^2 = \sum_{k = K + 1}^{\infty}  \dotp{f_0}{\psi_k}^2 \leq  \frac{1}{\lambda_{K + 1}(\Delta_P)^s} \sum_{k = K + 1}^{\infty} \lambda_{k + 1}(\Delta_P)^s \dotp{f_0}{\psi_k}^2 \leq \frac{\|f_0\|_{\mc{H}^s}}{\bigl[\lambda_{K + 1}(\Delta_P)\bigr]^s}.
+\end{equation*}
+On the other hand, the variance term can be written as the sum of the variance of each empirical Fourier coefficient,
+\begin{equation}
+\label{pf:spectral_series_estimation_2}
+\Ebb\| \wt{f} - \Ebb[\wt{f}]\|_P^2 = \sum_{k = 1}^{K} \Var\Bigl[\dotp{{\bf Y}}{\psi_k}_n\Bigr] 
+\end{equation}
+Then from the law of total variance,
+\begin{equation*}
+ \Var\bigl[\dotp{{\bf Y}}{\psi_k}_n\bigr] = \Var\Bigl[\Ebb[\dotp{Y}{\psi_k}_n|{\bf X}\Bigr] + \Ebb\Bigl[\Var[\dotp{Y}{\psi_k}_n|{\bf X}\Bigr] = \Var\Bigl[\dotp{f_0}{\psi_k}_n\Bigr] + \frac{1}{n}\Ebb\Bigl[\|\psi_k\|_n^2\Bigr].
+\end{equation*}
+Finally, $\Ebb \|\psi_k\|_n^2 = \|\psi_k\|_P^2 = 1$, 
+\begin{equation*}
+\Var\Bigl[\dotp{f_0}{\psi_k}_n\Bigr] = \frac{1}{n} \Var\Bigl[f_0(X)\psi_k(X)\Bigr] \leq \frac{1}{n} \Ebb\Bigl[\Bigl(f_0(X)\psi_k(X)\Bigr)^2\Bigr],
+\end{equation*}
+and plugging back into~\eqref{pf:spectral_series_estimation_2} yields the claim.
 
@@
 
-In Lemma~\textcolor{red}{(?)}, we show that the graph Sobolev semi-norm (V-statistic) decomposes into an unbiased estimate of the non-local Sobolev semi-norm (U-statistic), and a pure bias term.
-We also establish that the pure bias term will be small (in expectation) relative to the U-statistic whenever $\varepsilon$ is sufficiently small.
+In Lemma~\ref{lem:graph_seminorm_bias}, we decompose the graph Sobolev semi-norm (a V-statistic) into an unbiased estimate of the non-local Sobolev semi-norm (a U-statistic), and a pure bias term.
+We establish that the pure bias term will be small (in expectation) relative to the U-statistic whenever $\varepsilon$ is sufficiently small.
 \begin{lemma}
@@
 \end{equation*}
-In Lemma~\textcolor{red}{(?)}, we show that if $f \in H^1(\mc{X};M)$, then for any $i\bj \in [n]^{s + 1}$ which contains a total of $k + 1$ distinct indices, 
+In Lemma~\ref{lem:graph_seminorm_bias2}, we show that if $f \in H^1(\mc{X};M)$, then for any $i\bj \in [n]^{s + 1}$ which contains a total of $k + 1$ distinct indices, 
 \begin{equation*}
@@
 Note immediately that $G_2(x) = 0$ if $s = 1$.
-Otherwise if $s \geq 2$, then $q = s - 1$.Recalling that $|r_{x + z\varepsilon}^{s - 1}(x; p)| \leq C\varepsilon^{s - 1}\|p\|_{C^{s - 1}(\mc{X})}$ for any $z \in B(0,1)$, and that $d_x^jf(\cdot)$ is a $j$-homogeneous function, we have that
+Otherwise if $s \geq 2$, then $q = s - 1$.
+Recalling that $|r_{x + z\varepsilon}^{s - 1}(x; p)| \leq C\varepsilon^{s - 1}\|p\|_{C^{s - 1}(\mc{X})}$ for any $z \in B(0,1)$, and that $d_x^jf(\cdot)$ is a $j$-homogeneous function, we have that
 \begin{align}
@@
 First, we  recall that $L_{P,\varepsilon}$ is self-adjoint in $L^2(P)$, meaning $\dotp{L_{P,\varepsilon}f}{g}_{P} = \dotp{f}{L_{P,\varepsilon}g}_{P}$ for all $f, g \in L^2(P)$.
-We will also use the fact that when $s = 1$, the non-local seminorm is a \emph{non-local Dirichlet energy},
-\begin{equation*}
+We also recall the definition of the Dirichlet energy $E_{P,\varepsilon}(f;\mc{X})$,
+\begin{equation}
+\label{eqn:dirichlet_energy}
 \dotp{L_{P,\varepsilon}f}{f}_{P} = \frac{1}{\varepsilon^{d + 2}}\int_{\mc{X}} \int_{\mc{X}} \bigl(f(x) - f(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) =: E_{P,\varepsilon}(f;\mc{X}).
-\end{equation*}
+\end{equation}
 Finally, we recall a result of \textcolor{blue}{(Green 21)}: there exist constants $c_0$ and $C_0$ which do not depend on $M$, such that for all $\varepsilon < c_0$ and for any $f \in H^1(\mc{X};M)$,
@@
 \begin{equation*}
-\dotp{L_{P,\varepsilon}^sf}{f}_P = \dotp{L_{P,\varepsilon}^{k + 1}f}{L_{P,\varepsilon}^{k}f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}^{k};\mc{X}).
+\dotp{L_{P,\varepsilon}^sf}{f}_P = \dotp{L_{P,\varepsilon}^{k + 1}f}{L_{P,\varepsilon}^{k}f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\mc{X}).
 \end{equation*}
@@
 \begin{align}
-E_{P,\varepsilon}(L_{P,\varepsilon}^{k};\mc{X}) & = \frac{1}{\varepsilon^{d + 2}}\Biggl\{\int_{\mc{X}_{k\varepsilon}} \int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - L_{P,\varepsilon}^kf(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) \nonumber \\
+E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\mc{X}) & = \frac{1}{\varepsilon^{d + 2}}\Biggl\{\int_{\mc{X}_{k\varepsilon}} \int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - L_{P,\varepsilon}^kf(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) \nonumber \\
 & \quad + \int_{\partial_{k\varepsilon}\mc{X}} \int_{\partial_{k\varepsilon}\mc{X}} \bigl(L_{P,\varepsilon}^kf(x) - L_{P,\varepsilon}^kf(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x)\Biggr\} \nonumber \\
-& := E_{P,\varepsilon}(L_{P,\varepsilon}^{k};\mc{X}_{k\varepsilon}) + E_{P,\varepsilon}(L_{P,\varepsilon}^{k};\partial_{k\varepsilon}\mc{X}) \label{pf:estimate_nonlocal_seminorm_1.5}
+& := E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\mc{X}_{k\varepsilon}) + E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\partial_{k\varepsilon}\mc{X}) \label{pf:estimate_nonlocal_seminorm_1.5}
 \end{align}
@@
 \label{pf:estimate_nonlocal_seminorm_2}
-E_{P,\varepsilon}(L_{P,\varepsilon}^{k};\mc{X}_{k\varepsilon}) \leq 3 \sigma_{\eta}^{2k} E_{P,\varepsilon}(\Delta_P^kf;\mc{X}_{k\varepsilon}) + \frac{2}{\varepsilon^{d + 2}}\int_{\mc{X}_{k\varepsilon}} \int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_P^kf(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x).
+E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\mc{X}_{k\varepsilon}) \leq 3 \sigma_{\eta}^{2k} E_{P,\varepsilon}(\Delta_P^kf;\mc{X}_{k\varepsilon}) + \frac{2}{\varepsilon^{d + 2}}\int_{\mc{X}_{k\varepsilon}} \int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_P^kf(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x).
 \end{equation}
@@
 \begin{equation*}
-E_{P,\varepsilon}(L_{P,\varepsilon}^{k};\partial_{k\varepsilon}\mc{X}) \leq \frac{4p_{\max}^2}{\varepsilon^2}\|L_{P,\varepsilon}^kf\|_{L^2(\partial_{k\varepsilon}\mc{X})} \leq C M^2.
+E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\partial_{k\varepsilon}\mc{X}) \leq \frac{4p_{\max}^2}{\varepsilon^2}\|L_{P,\varepsilon}^kf\|_{L^2(\partial_{k\varepsilon}\mc{X})} \leq C M^2.
 \end{equation*}
@@
 	\begin{equation*}
-	\Ebb\bigl[|D_{\bj}f(X_i)| \cdot |f(X_i) - f(X_{\bj_1})|\bigr] \leq C \varepsilon^{2 + dk}M^2,
+	\Ebb\Bigl[|D_{\bj}f(X_i)| \cdot |f(X_i) - f(X_{\bj_1})|\Bigr] \leq C \varepsilon^{2 + dk}M^2,
 	\end{equation*}
@@
 
-\section{Graph quadratic form, manifold domain}
+\section{Graph Sobolev semi-norm, manifold domain}
 \label{sec:graph_quadratic_form_manifold}
-
-\paragraph{Discussion of requirements that $s \leq 3$.}
-Recall from our discussion in Section~\ref{subsec:analysis} that an essential step in the proof of Lemma~\textcolor{red}{(?)}---which establishes a high-probability upper bound on $\dotp{L_{n,\varepsilon}^sf_0}{f_0}_n$---is relating the iterated non-local operator $L_{P,\varepsilon}^j$ to the weighted Laplace-Beltrami operator $\Delta_P^j$, for $j = \floor{s/2}$.
-Suppose for the moment $s = 3$, and $f \in C^3(\mc{X};1)$ and $p$ is uniform.
-In this case $\floor{s/2} = 1$, and the relation between $L_{P,\varepsilon}$ to $\Delta_P$ has been worked out by \textcolor{red}{(Burago2013, GarciaTrillos2019, Calder2019, GarciaTrillos2020)}.
-This relation is accomplished by means of an intermediary nonlocal operator, as described by the following chain of estimates:
-\begin{align*}
-L_{P,\varepsilon} f(x) & = \frac{1}{\varepsilon^{m + 2}}\int_{\mc{X}} \bigl(f(z) - f(x)\bigr)\eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,d\mu(z) \\
-& \overset{(i)}{=} \frac{1}{\varepsilon^{m + 2}} \int_{\mc{X}} \bigl(f(z) - f(x)\bigr)\eta\biggl(\frac{d_{\mc{X}}(z,x)}{\varepsilon}\biggr) \,d\mu(z) + O(\varepsilon) \\
-& = \frac{1}{\varepsilon^{m + 2}} \int_{T_x(\mc{X})} \Bigl(f\bigl(\exp_x(v)\bigr) - f\bigl(\exp_x(0)\bigr)\Bigr)\eta\biggl(\frac{d_{\mc{X}}\bigl(\exp_x(v),\exp_x(0)\bigr)}{\varepsilon}\biggr) J_x(v) \,dv + O(\varepsilon) \\
-& \overset{(ii)}{=} \frac{1}{\varepsilon^{m + 2}} \int_{T_x(\mc{X})} \Bigl(f\bigl(\exp_x(v)\bigr) - f\bigl(\exp_x(0)\bigr)\Bigr)\eta\biggl(\frac{d_{\mc{X}}\bigl(\exp_x(v),\exp_x(0)\bigr)}{\varepsilon}\biggr) \,dv + O(\varepsilon) \\
-& = \frac{1}{\varepsilon^{m + 2}} \int_{T_x(\mc{X})} \Bigl(f\bigl(\exp_x(v)\bigr) - f\bigl(\exp_x(0)\bigr)\Bigr)\eta\biggl(\frac{\|v\|}{\varepsilon}\biggr) \,dv + O(\varepsilon) \\
-& \overset{(iii)}{=} \Delta_Pf(x) + O(\varepsilon)
-\end{align*}
-Here, $d_{\mc{X}}(\cdot,\cdot)$ is the geodesic distance on $\mc{X}$, $J_x(v)$ denotes the Jacobian of the exponential map $\exp_x$ evaluated at $v \in T_x(\mc{M})$, and $h = f \circ \exp_x$.
-Then $(i)$ follows because $\bigl|d_{\mc{X}}(x,z)  - \|z - x\|\bigr| = O(\varepsilon^3)$ and $\eta$ is a Lipschitz function with bounded Lipschitz constant, $(ii)$ follows because the Jacobian $J_x(v)$ satisfies $|J_x(v)| = 1 + O(\varepsilon^2)$ for all $v \in B(0,\varepsilon) \subset T_x(\mc{X})$, and $(iii)$ follows since the function $h$ admits the Taylor expansion
-\begin{equation*}
-h(v) = h(0) + \dotp{\nabla h(0)}{v} + \frac{1}{2} \dotp{\nabla^2 h(0) v}{v} + O(\varepsilon^3),~~\textrm{for all $v \in B(0,\varepsilon) \subset T_x(\mc{X})$.}
-\end{equation*}
-Then, using the upper bound on $\max_{i = 1,\ldots,n}|L_{n,\varepsilon}f(X_i) - L_{P,\varepsilon}(X_i)| = O_\Pbb((n\varepsilon^{m + 4})^{-1})$ given in \textcolor{red}{(?)}, and the fact that $\Delta_Pf \in C^1(\mc{X};1)$, we have that,
-\begin{align*}
-\dotp{L_{n,\varepsilon}^3f}{f}_n & = \frac{1}{n^2 \varepsilon^{m + 2}} \sum_{i,j = 1}^{n} \bigl(L_{n,\varepsilon}f(X_i) - L_{n,\varepsilon}f(X_j)\bigr)^2 \eta\biggl(\frac{\|X_i - X_j\|}{\varepsilon}\biggr) \\
-& = \frac{1}{n^2 \varepsilon^{m + 2}} \sum_{i,j = 1}^{n} \bigl(L_{P,\varepsilon}f(X_i) - L_{P,\varepsilon}f(X_j)\bigr)^2 \eta\biggl(\frac{\|X_i - X_j\|}{\varepsilon}\biggr) + O_{\Pbb}\biggl(\frac{1}{n\varepsilon^{m + 4}}\biggr) \\
-& = \frac{1}{n^2 \varepsilon^{m + 2}} \sum_{i,j = 1}^{n} \bigl(\Delta_Pf(X_i) - \Delta_Pf(X_j)\bigr)^2 \eta\biggl(\frac{\|X_i - X_j\|}{\varepsilon}\biggr) + O_{\Pbb}\biggl(\frac{1}{n\varepsilon^{m + 4}}\biggr) + O_{\Pbb}(1) \\
-& = O_{\Pbb}\biggl(\frac{1}{n\varepsilon^{m + 4}}\biggr) + O_{\Pbb}(1) = O_{\Pbb}(1),
-\end{align*}
-with the last equality following when $\varepsilon = \Omega(n^{-1/(m + 4)})$.
-
-On the other hand when $s = 4$, and thus $j = 2$, we have only that
-\begin{equation*}
-L_{P,\varepsilon}^2f(x) = L_{P,\varepsilon}(\Delta_Pf + O(\varepsilon)) = \Delta_P^2f(x) + O(1/\varepsilon),
-\end{equation*}
-which is useless when $\varepsilon \to 0$.
-The bottom line is that in the manifold case, the $O(\varepsilon^3)$ error incurred by going between Euclidean norm and geodesic distance is negligible only when $j = 1$.
+In this section we prove Proposition~\ref{prop:graph_seminorm_manifold}.
+Note that when $s = 1$, the upper bound~\eqref{eqn:graph_seminorm_manifold} follows immediately from Lemma~\ref{lem:dirichlet_energy_sobolev} and Markov's inequality.
+
+On the other hand when $s = 2$ or $s = 3$, we prove Proposition~\ref{prop:graph_seminorm_manifold} by first establishing some intermediate results, many of which are analogous to results we have already shown in the flat Euclidean case.
+Indeed, in some ways the proof will be simpler in the manifold setting than in the flat Euclidean case: there is no boundary, and we do not need to analyze the iterated nonlocal Laplacian $L_{P,\varepsilon}^j$ for $j > 1$.
+
+That being said, as mentioned in our main text, in the manifold setting there is some extra error induced by using Euclidean rather than geodesic distance.
+We upper bound this error by comparing $L_{P,\varepsilon}$ to an alternative nonlocal Laplacian $\wt{L}_{P,\varepsilon}$, which is defined with respect to geodesic distance.
+Precisely, let $d_{\mc{X}}(x,x')$ denote the geodesic distance between $x,x' \in \mc{X}$, and define
+\begin{equation*}
+\wt{L}_{P,\varepsilon}f(x) := \int_{\mc{X}} \bigl(f(x') - f(x)\bigr) \eta \biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr) p(x') \,dx'.
+\end{equation*}
+
+We show the following results, each of which hold under the same assumptions as Proposition~\ref{prop:graph_seminorm_manifold}.
+\begin{itemize}
+	\item In Section~\ref{subsec:manifold_decomposition_graph_seminorm} we show that the graph Sobolev seminorm $\dotp{L_{n,\varepsilon}^sf}{f}_n$ is upper bounded by the sum of a nonlocal seminorm and a pure bias term: specifically, with probability at least $1 - 2\delta$,
+	\begin{equation}
+	\label{pf:graph_seminorm_manifold_1}
+	\dotp{L_{n,\varepsilon}^sf}{f}_n \leq \frac{\dotp{L_{P,\varepsilon}^sf}{f}_P}{\delta} + C_1\frac{\varepsilon^2}{n\varepsilon^{2s + m}}M^2.
+	\end{equation}
+	This upper bound is essentially the same as~\eqref{pf:graph_seminorm_ho_1}, but with the intrinsic dimension $m$ taking the place of the ambient dimension $d$.
+The pure bias term will be of at most constant order when $\varepsilon \gtrsim n^{-1/(2(s-1) + m)}$.
+	\item In Section~\ref{subsec:error_euclidean_distance}, we show that the error incurred by using the ``wrong'' metric is negligible. Precisely, we find that
+	\begin{equation}
+	\label{eqn:nonlocal_laplacian_geodesic_error}
+	\|L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f\|_{L^2(\mc{X})}^2 \leq C_2 \varepsilon^2 |f|_{H^1(\mc{X})}^2.
+	\end{equation}
+	\item In Section~\ref{subsec:manifold_approximation_error_nonlocal_laplacian}, we analyze the approximation error of $\wt{L}_{P,\varepsilon}$.
+We show that when $f \in H^2(\mc{X})$ and $p \in C^1(\mc{X})$,
+	\begin{equation}
+	\label{eqn:nonlocal_laplacian_approximation_error_manifold_l2}
+	\|\wt{L}_{P,\varepsilon}f\|_{\Leb^2(\mc{X})}^2 \leq C_3 \|f\|_{H^2(\mc{X})}^2,
+	\end{equation}
+	whereas if $f \in H^3(\mc{X})$ and $p \in C^2(\mc{X})$, 
+	\begin{equation}
+	\label{eqn:nonlocal_laplacian_approximation_error_manifold_sobolev}
+	\|\wt{L}_{P,\varepsilon}f - \sigma_{\eta}\Delta_Pf\|_{\Leb^2(\mc{X})}^2 \leq C_3 \varepsilon^2 \|f\|_{H^3(\mc{X})}^2.
+	\end{equation}
+	\item In Section~\ref{subsec:manifold_estimate_nonlocal_seminorm}, we use the results of the preceding two sections to show that if $f \in H^s(\mc{X})$ and $p \in C^{s - 1}(\mc{X})$, then 
+	\begin{equation}
+	\label{eqn:manifold_nonlocal_seminorm}
+	\dotp{L_{P,\varepsilon}^sf}{f}_P \leq C_4\|f\|_{H^s(\mc{X})}^2.
+	\end{equation}
+	\item In Section~\ref{subsec:manifold_integrals} we state some technical results used in the previous sections.
+\end{itemize}
+We point out that when $f$ is H\"{o}lder smooth, results analogous to~\eqref{eqn:nonlocal_laplacian_approximation_error_manifold_sobolev} have been established in \citet{calder2019}.
+When $f$ is Sobolev smooth, our analysis (which relies heavily on Taylor expansions) is largely similar, except that the remainder term in the relevant Taylor expansion will be bounded in $L^2(\mc{X})$ norm rather than $L^{\infty}(\mc{X})$ norm.
+This is analogous to the situation in the flat Euclidean model.
+
+\paragraph{Proof (of Proposition~\ref{prop:graph_seminorm_manifold}).} Follows immediately from~\eqref{pf:graph_seminorm_manifold_1} and~\eqref{eqn:manifold_nonlocal_seminorm}. \qed
+
+\subsection{Decomposition of graph Sobolev seminorm}
+\label{subsec:manifold_decomposition_graph_seminorm}
+The proof of~\eqref{pf:graph_seminorm_manifold_1} is identical to the proof of~\eqref{pf:graph_seminorm_ho_1}, except substituting the intrinsic dimension $m$ for ambient dimension $d$, and using Lemma~\ref{lem:manifold_graph_seminorm_bias2} rather than Lemma~\ref{lem:graph_seminorm_bias2}.
+
+\subsection{Error due to Euclidean Distance}
+\label{subsec:error_euclidean_distance}
+
+In this section, we prove~\eqref{eqn:nonlocal_laplacian_geodesic_error}. By applying Cauchy-Schwarz we obtain an upper bound on $|L_{P,\varepsilon}f(x) - \wt{L}_{P,\varepsilon}f(x)|^2$:
+\begin{align}
+\bigl[L_{P,\varepsilon}f(x) - \wt{L}_{P,\varepsilon}f(x)\bigr]^2 & \leq \frac{p_{\max}^2}{\varepsilon^{2(2 + m)}} \int_{\mc{X}} \bigl[f(x') - f(x)\bigr]^2 \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') \nonumber \\
+& \quad \cdot \int_{\mc{X}} \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') \nonumber \\
+& = \frac{1}{\varepsilon^{2(2 + m)}} A_1(x) \cdot A_2(x) \label{pf:error_euclidean_distance_0}
+\end{align} 
+Thus we have upper bounded $|L_{P,\varepsilon}f(x) - \wt{L}_{P,\varepsilon}f(x)|^2$ by the product of two terms, each of which we now suitably bound.
+
+To do so, we will use the following estimates, from Proposition 4 of \cite{trillos2019}: letting $R$ denote the reach of $\mc{X}$, for all $\|x' - x\| \leq R/2$,
+\begin{equation}
+\label{eqn:distance_error}
+\|x' - x\| \leq d_{\mc{X}}(x',x) \leq \|x' - x\| + \frac{8}{R^2} \|x' - x\|^3.
+\end{equation}
+\paragraph{Upper bound on $A_1(x)$.}
+From here forward we will assume $\varepsilon < R/2$.
+Consequently $\eta(\|x' - x\|/\varepsilon) \geq \eta(d_{\mc{X}}(x',x)/\varepsilon)$.
+Furthermore, letting $L_{\eta}$ denote the Lipschitz constant of $\eta$, and setting $\wt{\varepsilon} := (1 + 27\varepsilon^2/R^2)\varepsilon$ we have that
+\begin{equation*}
+\biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \leq \frac{L_{\eta} 8 \varepsilon^2}{R^2} \cdot \1\bigl\{d_{\mc{X}}(x',x) \leq \varepsilon\bigr\} + \|\eta\|_{\infty} \cdot 1\{\varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon}\}.
+\end{equation*}
+Thus,
+\begin{equation*}
+A_1(x) \leq \frac{8L_{\eta}\varepsilon^2}{R^2}\int_{\mc{X}}\bigl[f(x') - f(x)\bigr]^2 \1\{\|x' - x\| \leq \varepsilon\} \,d\mu(x') + \|\eta\|_{\infty} \int_{\mc{X}}\bigl[f(x') - f(x)\bigr]^2 \1\bigl\{\varepsilon \|x' - x\| \leq \wt{\varepsilon}\bigr\} \,d\mu(x') \\
+\end{equation*}
+Integrating over $\mc{X}$, we conclude from Lemma~\ref{lem:dirichlet_energy_remainder} and Lemma~3.3 of \citep{burago2014} and  that
+\begin{equation*}
+\int_{\mc{X}} A_1(x) \,d\mu(x) \leq \frac{8L_{\eta}\nu_m\varepsilon^2}{R^2(m + 2)} \Bigl(1 + CmKR^2\Bigr) \varepsilon^{m + 2} |f|_{H^1(\mc{X})}^2 + C\|\eta\|_{\infty}\varepsilon^{m + 4} |f|_{H^1(\mc{X})}^2 =: C_5 \varepsilon^{m + 4}|f|_{H^1(\mc{X})}.
+\end{equation*}
+
+\paragraph{Upper bound on $A_2(x)$.}
+Integrating over $x' \in \mc{X}$, we see that
+\begin{align}
+\int_{\mc{X}} \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') & \leq \frac{8L_{\eta}\varepsilon^2}{R^2} \int_{\mc{X}} \1\bigl\{d_{\mc{X}}(x',x)\bigr\} \,d\mu(x') + p_{\max}\|\eta\|_{\infty} \int_{\mc{X}} \1\bigl\{ \varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon} \bigr\} \,d\mu(x') \nonumber \\
+& = \frac{8L_{\eta}\varepsilon^2}{R^2} \cdot \mu\bigl(B(x,\varepsilon)\bigr) +  p_{\max}\|\eta\|_{\infty} \Bigl[\mu\bigl(B(x,\wt{\varepsilon})\bigr) - \mu\bigl(B(x,\varepsilon)\bigr) \Bigr].
+\label{pf:error_euclidean_distance_1}
+\end{align}
+Equation (1.36) in \cite{trillos2019} states that
+\begin{equation*}
+\bigl|\mu(B_\mc{X}(x,\varepsilon))  - \omega_m \varepsilon^m\bigr|  \leq CmK\varepsilon^{m + 2},
+\end{equation*}
+where $K$ is an upper bound on the \textcolor{red}{sectional curvature} of $\mc{X}$. Plugging this back into~\eqref{pf:error_euclidean_distance_1}, we conclude that
+\begin{align*}
+\int_{\mc{X}} \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') & \leq  \frac{8L_{\eta}\varepsilon^2}{R^2}\Bigl[\omega_m\varepsilon^{m} + CmK\varepsilon^{m + 2}\Bigr] + \|\eta\|_{\infty} \Bigl[\omega_m(\wt{\varepsilon}^m - {\varepsilon}^m) + 2CmK\varepsilon^{m + 2}\Bigr] \\
+& \leq \frac{8L_{\eta}\varepsilon^2}{R^2}\Bigl[\omega_m\varepsilon^{m} + R^2CmK\varepsilon^{m}\Bigr] + \|\eta\|_{\infty} \varepsilon^{m + 2}\Bigl[\frac{27\omega_m}{R^2} + 2CmK\Bigr] \\
+& =: C_6 \varepsilon^{m + 2}.
+\end{align*}
+
+\paragraph{Putting together the pieces.}
+Plugging our upper bounds on $A_1(x)$ and $A_2(x)$ back into~\eqref{pf:error_euclidean_distance_0}, we deduce that
+\begin{align*}
+\|\wt{L}_{P,\varepsilon}f - L_{P,\varepsilon}f\|_{L^2(\mc{X})}^2 & \leq \frac{1}{\varepsilon^{2(2 + m)}} \int_{\mc{X}} A_1(x) \cdot A_2(x) \,d\mu(x) \\
+& \leq \frac{C_6}{\varepsilon^{(2 + m)}} \int_{\mc{X}} A_1(x) \,d\mu(x) \\
+& \leq C_5 C_6 \varepsilon^2 |f|_{H^1(\mc{X})}^2,
+\end{align*}
+thus proving the claimed result.
+
+\subsection{Approximation Error of non-local Laplacian}
+\label{subsec:manifold_approximation_error_nonlocal_laplacian}
+Fix $x \in \mc{X}$.
+We begin with a pointwise estimate of $\wt{L}_{P,\varepsilon}f$, facilitated by expressing $w(v) = f(\exp_x(v))$ and $q(v) = p(\exp_x(v))$ in normal coordinates, as in \citep{calder2019}.
+Letting $J_x(v)$ be the Jacobian of the exponential map $\exp_x: B(0,\varepsilon) \subseteq T_x(\mc{X}) \to B_{\mc{X}}(x,\varepsilon)$ evaluated at $v \in T_x(\mc{X})$, we have
+\begin{align*}
+\wt{L}_{P,\varepsilon}f(x) & = \frac{1}{\varepsilon^{m + 2}} \int_{\mc{X}} \bigl(f(x') - f(x)\bigr) \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr) \,dP(x') \\
+& = \frac{1}{\varepsilon^{m + 2}} \int_{B(0,\varepsilon) \subset T_x(\mc{X})} \bigl(w(v) - w(0)\bigr) \eta\biggl(\frac{\|v\|}{\varepsilon}\biggr) J_x(v) q(v) \,dv \\
+& = \frac{1}{\varepsilon^{2}}\biggl\{\int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr) \eta(\|v\|) q(\varepsilon v) \,dv + \int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr) \eta(\|v\|) q(\varepsilon v) \bigl(J_x(\varepsilon v) - 1\bigr) \,dv \biggr\} \\
+& = A_1(x) + A_2(x)
+\end{align*}
+Note that $w$ and $q$ have the same smoothness properties as $f$ and $p$.
+Moreover, arguing exactly as we did in the flat Euclidean case, we can show that when $f \in H^2(\mc{X})$ and $p \in C^1(\mc{X})$, then
+\begin{equation*}
+\|A_1\|_{\Leb^2(\mc{X})}^2 \leq C \|f\|_{H^2(\mc{X})}^2
+\end{equation*}
+whereas if $f \in H^3(\mc{X})$ and $p \in C^2(\mc{X})$ then 
+\begin{equation*}
+\|A_1 - \sigma_{\eta} \Delta_Pf\|_{\Leb^2(\mc{X})}^2 \leq C\|f\|_{H^3(\mc{X})}^2 \varepsilon^2.
+\end{equation*}
+\textcolor{red}{(TODO): This is correct, but feels unacceptably vague.
+On the other hand, I don't want to have to redo the entire proof from the flat Euclidean section, since all the steps will be basically identical.
+Ask Ryan + Siva for guidance.}
+
+Therefore it remains only to upper bound $A_2$ in $\Leb^2(\mc{X})$ norm.
+To do so, we recall (1.34) of \cite{trillos2019}: for any $\varepsilon < i_0$ and all $x \in \mc{X}$, the Jacobian $J_x(v)$ satisfies the upper bound
+\begin{equation*}
+|J_x(v) - 1| \leq CmK\varepsilon^2, \quad  \textrm{for all} ~~ v \in B(0,\varepsilon) \subseteq T_x(\mc{X}).
+\end{equation*}
+Combining this estimate with the Cauchy-Schwarz inequality, we conclude that
+\begin{align*}
+\|A_2\|_{\Leb^2(\mc{X})}^2 & \leq Cm^2K^2 \biggl[\int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr)^2 \eta(\|v\|) q(\varepsilon v) \,dv\biggr] \cdot \biggl[\int_{B(0,1)} \eta(\|v\|) q(\varepsilon v) \,dv\biggr] \\
+& \leq Cm^2K^2 \sigma_{\eta} (1 + L_q\varepsilon)  \int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr)^2 \eta(\|v\|) q(\varepsilon v) \,dv \\
+& \leq Cm^2K^2 \sigma_{\eta}^2 (1 + L_q\varepsilon) p_{\max}  \varepsilon^2 |f|_{H^1(\mc{X})}^2,
+\end{align*}
+with the final inequality following from (3.2) of~\cite{burago2014}. Combining our estimates on $A_1$ and $A_2$ yields the claim.
+
+\subsection{Estimate of non-local Sobolev seminorm}
+\label{subsec:manifold_estimate_nonlocal_seminorm}
+In this subsection we establish that the upper bound~\eqref{eqn:manifold_nonlocal_seminorm} holds when $f \in H^s(\mc{X})$ and $p \in C^{s - 1}(\mc{X})$. We first consider $s = 2$, and then $s = 3$.
+
+\paragraph{Case 1: $s = 2$.}
+When $s = 2$, the triangle inequality implies that
+\begin{equation*}
+\dotp{L_{P,\varepsilon}^sf}{f}_P \leq 2p_{\max}\Bigl(\|L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}\|_{L^2(\mc{X})}^2 + \|\wt{L}_{P,\varepsilon}f\|_{L^2(\mc{X})}^2\Bigr)
+\end{equation*}
+The first term on the right hand side is upper bounded in~\eqref{eqn:nonlocal_laplacian_geodesic_error}, and the second term is upper bounded in~\eqref{eqn:nonlocal_laplacian_approximation_error_manifold_l2}.
+Together these estimates imply the claim.
+
+\paragraph{Case 2: $s = 3$.}
+When $s = 3$, the triangle inequality implies that
+\begin{equation*}
+\dotp{L_{P,\varepsilon}^sf}{f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}f;\mc{X}) \leq 3\Bigl( E_{P,\varepsilon}(L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f;\mc{X}) +  E_{P,\varepsilon}(\wt{L}_{P,\varepsilon}f - \sigma_{\eta} \Delta_Pf;\mc{X}) + \sigma_{\eta}^2 E_{P,\varepsilon}(\Delta_Pf;\mc{X})\Bigr)
+\end{equation*}
+We now upper bound each of the three terms on the right hand side of the above inequality.
+First, we note that by Lemma~\ref{lem:dirichlet_energy_l2} and~\eqref{eqn:nonlocal_laplacian_geodesic_error},
+\begin{equation*}
+ E_{P,\varepsilon}(L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f;\mc{X}) \leq  \frac{C}{\varepsilon^2}\|L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f\|_{\Leb^2(\mc{X})}^2 \leq C |f|_{H^1(\mc{X})}^2.
+\end{equation*}
+An equivalent upper bound on $E_{P,\varepsilon}(\wt{L}_{P,\varepsilon}f - \sigma_{\eta} \Delta_Pf;\mc{X})$ follows from Lemma~\ref{lem:dirichlet_energy_l2} and~\eqref{eqn:nonlocal_laplacian_approximation_error_manifold_sobolev}.
+Finally, we notice that $f \in H^3(\mc{X})$ and $p \in C^2(\mc{X})$ implies $\Delta_Pf \in H^1(\mc{X})$, and furthermore $|\Delta_Pf|_{H^1(\mc{X})} \leq \|p\|_{C^2(\mc{X})} \cdot \|f\|_{H^3(\mc{X})}$.
+We conclude from Lemma~\ref{lem:dirichlet_energy_sobolev} that
+\begin{equation*}
+E_{P,\varepsilon}(\Delta_Pf;\mc{X}) \leq C |\Delta_Pf|_{H^1(\mc{X})}^2 \leq C \|f\|_{H^3(\mc{X})}^2,
+\end{equation*}
+where in the final inequality we have absorbed $\|p\|_{C^2(\mc{X})}$ into the constant $C$. Together, these upper bounds prove the claim.
+
+\subsection{Integrals}
+\label{subsec:manifold_integrals}
+Recall the Dirichlet energy $E_{P,\varepsilon}(f;\mc{X}) = \dotp{L_{P,\varepsilon}f}{f}_P$, defined in~\eqref{eqn:dirichlet_energy}.
+Now we establish some estimates on $E_{P,\varepsilon}(f;\mc{X})$ under Model~\ref{def:model_manifold}, and under various assumptions regarding the regularity of $f$.
+\begin{lemma}
+	\label{lem:dirichlet_energy_l2}
+	Suppose Model~\ref{def:model_manifold}, and additionally that $f \in L^2(\mc{X})$. Then there exists a constant $C$ such that
+	\begin{equation}
+	\label{eqn:dirichlet_energy_l2}
+	E_{P,\varepsilon}(f;\mc{X}) \leq \frac{C}{\varepsilon^2} \|f\|_{L^2(\mc{X})}^2.
+	\end{equation}
+\end{lemma}
+\begin{lemma}
+	\label{lem:dirichlet_energy_sobolev}
+	Suppose Model~\ref{def:model_manifold}, and additionally that $f \in H^1(\mc{X})$. Then there exist constants $c$ and $C$ which do not depend on $f$ such that for any $0 < \varepsilon < c$,
+	\begin{equation}
+	\label{eqn:dirichlet_energy_sobolev}
+	E_{P,\varepsilon}(f;\mc{X}) \leq C |f|_{H^1(\mc{X})}^2.
+	\end{equation}
+\end{lemma}
+
+We use Lemma~\ref{lem:dirichlet_energy_remainder} to help upper bound the error incurred by using $\|\cdot\|$ rather than $d_{\mc{X}}(\cdot,\cdot)$.
+Recall the notation $\wt{\varepsilon} = (1 + 27\varepsilon^2/R^2)\varepsilon$, where $R$ is the reach of $\mc{X}$.
+\begin{lemma}
+	\label{lem:dirichlet_energy_remainder}
+	Suppose Model~\ref{def:model_manifold}, and additionally that $f \in H^1(\mc{X})$. There exist constants $c$ and $C$ such that for any $\varepsilon < c$,
+	\begin{equation*}
+	\int_{\mc{X}} \int_{\mc{X}} \bigl(f(x') - f(x)\bigr)^2 \1\{\varepsilon < \|x' - x\| < \wt{\varepsilon}\} \,dP(x') \,dP(x) \leq C \varepsilon^{4 + m} \|f\|_{H^1(\mc{X})}^2
+	\end{equation*}
+\end{lemma}
+
+Finally, we use Lemma~\ref{lem:manifold_graph_seminorm_bias2} to show that the pure bias component of $\dotp{L_n^sf,f}_n$ is small in expectation.
+This is analogous to Lemma~\ref{lem:graph_seminorm_bias2}, except assuming Model~\ref{def:model_manifold} rather than Model~\ref{def:model_flat_euclidean}.
+\begin{lemma}
+	\label{lem:manifold_graph_seminorm_bias2}
+	Assume Model~\ref{def:model_manifold}.
+Suppose $f \in H^1(\mc{X})$, and let $D_if$ be defined with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_manifold}.
+Then there exists a constant $C$ which does not depend on $f$ or $n$, such that for any $i \in [n]$ and $\bj \in [n]^s$,
+	\begin{equation*}
+	\Ebb\Bigl[|D_{\bj}f(X_i)| \cdot |f(X_i) - f(X_{\bj_1})|\Bigr] \leq C \varepsilon^{2 + mk} \cdot \|f\|_{H^1(\mc{X})}^2,
+	\end{equation*}
+	where $k + 1$ is the number of distinct indices in $i\bj$. 
+\end{lemma}
+
+\paragraph{Proof (of Lemmas~\ref{lem:dirichlet_energy_l2} and~\ref{lem:dirichlet_energy_sobolev}).}
+Define the non-local energy $\wt{E}_{P,\varepsilon}$ with respect to geodesic distance,
+\begin{equation*}
+\wt{E}_{P,\varepsilon}(f;{\mc{X}}) := \dotp{\wt{L}_{P,\varepsilon}f}{f}_{P} = \int_{\mc{X}} \int_{\mc{X}} \bigl(f(x') - f(x)\bigr)^2 \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr) \,dP(x') \,dP(x).
+\end{equation*}
+From the lower bound in~\eqref{eqn:distance_error}, it follows that $E_{P,\varepsilon}(f;X) \leq \wt{E}_{P,\varepsilon}(f;{\mc{X}})$, and from the upper bounds $p(x) \leq p_{\max}$ and $\eta(|x|) \leq \|\eta\|_{\infty} \cdot \1{x \in [-1,1]}$ we further have
+\begin{equation*}
+\wt{E}_{P,\varepsilon}(f;{\mc{X}}) \leq p_{\max}^2 \|\eta\|_{\infty} \cdot \int_{\mc{X}} \int_{B_{\mc{X}}(\varepsilon)} \bigl(f(x') - f(x)\bigr)^2 \,d\mu(x') \,d\mu(x).
+\end{equation*}
+The estimates~\eqref{eqn:dirichlet_energy_l2} and~\eqref{eqn:dirichlet_energy_sobolev} then respectively follow from (3.1) and Lemma~3.3 of \cite{burago2014}.
+
+\paragraph{Proof (of Lemma~\ref{lem:dirichlet_energy_remainder}).}
+\textcolor{red}{(TODO)}
+
+\paragraph{Proof (of Lemma~\ref{lem:manifold_graph_seminorm_bias2}).}
+The proof of Lemma~\ref{lem:manifold_graph_seminorm_bias2} is identical to the proof of Lemma~\ref{lem:graph_seminorm_bias2}, upon substituting the ambient dimension $m$ for the intrinsic dimension $d$, and using Lemma~\ref{lem:dirichlet_energy_sobolev} rather than Lemma~\ref{lem:dirichlet_estimate_nonlocal_laplacian} to establish~\eqref{pf:graph_seminorm_bias2_1}.
 
@@
 In this Section we prove Proposition~\ref{prop:empirical_norm_sobolev} (in Section~\ref{subsec:empirical_norm_sobolev}).
-We also prove an analogous result when $\mc{X}$ is a manifold as in Model~\ref{def:model_manifold} (in Section~\ref{subsec:empirical_norm_sobolev_manifold}), and we prove some sharper estimates in the case where the function is bounded in H\"{o}lder norm, in Section~\ref{subsec:empirical_norm_holder}.
+We also prove an analogous result when $\mc{X}$ is a manifold as in Model~\ref{def:model_manifold} (in Section~\ref{subsec:empirical_norm_sobolev_manifold}).
 
@@
 \label{subsec:empirical_norm_sobolev}
-\textcolor{red}{(TODO)}
-
-\subsection{Analysis of empirical norm in manifold setting}
-\label{subsec:empirical_norm_soboolev_manifold}
-
-\subsection{Better bounds assuming H\"{o}lder smoothness}
-\label{subsec:empirical_norm_holder}
+In this section we establish Proposition~\ref{prop:empirical_norm_sobolev}.
+As mentioned, the proof of this Proposition follows from the Gagliardo-Nirenberg interpolation inequality, and a one-sided Bernstein's inequality (Lemma~\ref{lem:one_sided_bernstein}).
+
+\begin{lemma}[Gagliardo-Nirenberg inequality]
+	\label{lem:gagliardo_nirenberg}
+	Suppose Model~\ref{def:model_flat_euclidean}, and that $f \in H^s(\mc{X})$ for some $s \geq d/4$. Then there exist constants $C_1$ and $C_2$ that do not depend on $f$, such that
+	\begin{equation}
+	\label{eqn:gagliardo_nirenberg}
+	\|f\|_{\Leb^4(\mc{X})} \leq C_1 |f|_{H^s(\mc{X})}^{d/4s} \|f\|_{\Leb^2(\mc{X})}^{1 - d/(4s)} + C_2 \|f\|_{\Leb^2(\mc{X})}
+	\end{equation}
+\end{lemma}
+\textcolor{red}{(TODO): Find the exact statement of this Lemma in \cite{leoni2017}.}
+
+\paragraph{Proof (of Proposition~\ref{prop:empirical_norm_sobolev}).}
+Rearranging~\eqref{eqn:gagliardo_nirenberg} and raising both sides to the $4$th power, we see that
+\begin{equation*}
+\frac{\Ebb[f^4(X)]}{\|f\|_P^4} \leq C \biggl(\frac{\|f\|_{\Leb^4(\mc{X})}}{\|f\|_{\Leb^2(\mc{X})}}\biggr)^4 \leq C_1\biggl(\frac{|f|_{H^s(\mc{X})}}{\|f\|_{\Leb^2(\mc{X})}}\biggr)^{d/s} + C_2,
+\end{equation*}
+here the constants $C_1,C_2$ are not the same as in~\eqref{eqn:gagliardo_nirenberg}.
+Therefore taking the constant $C$ in assumption~\eqref{eqn:empirical_norm_sobolev_1} to be sufficiently large relative to $C_1$ and $C_2$, we have that
+\begin{equation*}
+C_1\biggl(\frac{|f|_{H^s(\mc{X})}}{\|f\|_{\Leb^2(\mc{X})}}\biggr)^{d/s} \leq \frac{\delta n}{64},
+\end{equation*} 
+and consequently 
+\begin{equation*}
+\frac{\Ebb[f^4(X)]}{\|f\|_P^4} \leq \frac{\delta n}{8} + 8C_2^3.
+\end{equation*}
+The claim then follows from Lemma~\ref{lem:one_sided_bernstein}, upon taking $c = 1/(64C_2^3)$ in the statement of Proposition~\ref{prop:empirical_norm_sobolev}.
+
+\subsection{Proof of Proposition~\ref{prop:empirical_norm_sobolev_manifold}}
+\label{subsec:empirical_norm_sobolev_manifold}
+The proof of Proposition~\ref{prop:empirical_norm_sobolev_manifold} follows exactly the same steps as the proof of Proposition~\ref{prop:empirical_norm_sobolev}, except Lemma~\ref{lem:gagliardo_nirenberg} is replaced by Lemma~\ref{lem:gagliardo_nirenberg_manifold}.
+\begin{lemma}[(c.f Theorem~3.70 of~\citet{aubin2012})]
+	\label{lem:gagliardo_nirenberg_manifold}
+	Suppose Model~\ref{def:model_manifold}, and that $f \in H^s(\mc{X})$ for some $s \geq m/4$. Then there exist constants $C_1$ and $C_2$ that do not depend on $f$, such that
+	\begin{equation}
+	\label{eqn:gagliardo_nirenberg_manifold}
+	\|f\|_{\Leb^4(\mc{X})} \leq C_1 |f|_{H^s(\mc{X})}^{m/4s} \|f\|_{\Leb^2(\mc{X})}^{1 - m/(4s)} + C_2 \|f\|_{\Leb^2(\mc{X})}.
+	\end{equation}
+\end{lemma}
+
+\section{Proofs of Main Results}
+\label{sec:proofs_main_results}
+
+\subsection{Estimation Results}
+
+\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}.}
+We condition on the event that the design points $X_1,\ldots,X_n$ satisfy
+\begin{equation}
+\label{pf:laplacian_eigenmaps_estimation_fo_1}
+\dotp{L_{n,\varepsilon}f_0}{f_0}_n \leq \frac{C}{\delta}M^2 \quad \textrm{and} \quad \lambda_k \geq \min\{\lambda_k(\Delta_P), \varepsilon^{-2}\}~~\textrm{for all $2 \leq k \leq n$.}
+\end{equation}
+Note that by Propositions~\ref{prop:graph_seminorm_fo} and~\ref{prop:graph_eigenvalue}, these statements are both satisfied with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\}$.
+Conditional on~\eqref{pf:laplacian_eigenmaps_estimation_fo_1}, we have from Lemma~\ref{lem:fixed_graph_estimation} that for any $1 \leq K \leq n - 1$, with probability at least $1 - \exp(-K)$,
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 \leq C\biggl\{\frac{M^2}{\delta (\lambda_{K + 1}(\Delta_P) \wedge \varepsilon^{-2})} + \frac{K}{n}\biggr\}.
+\end{equation*}
+Further, from the bounds $\varepsilon \leq c_0 K^{-1/d}$ (Assumption~\ref{asmp:parameters_estimation_fo}) and $\lambda_{K + 1}(\Delta_P) \geq c K^{2/d}$ (Weyl's Law) we can simply the above expression the following,
+\begin{equation}
+\label{pf:laplacian_eigenmaps_estimation_fo_2}
+\|\wh{f} - f_0\|_n^2 \leq C\biggl\{\frac{M^2}{\delta}K^{-2/d} + \frac{K}{n}\biggr\}.
+\end{equation}
+We now upper bound the right hand side of~\eqref{pf:laplacian_eigenmaps_estimation_fo_2}, based on the value of $K$ chosen in~\ref{asmp:parameters_estimation_fo}.
+The classical case is $K = (M^2n)^{d/(2 + d)}$, in which case~\eqref{pf:laplacian_eigenmaps_estimation_fo_2} implies
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 \leq \frac{C}{\delta} M^2 (M^2n)^{-2/(2 + d)}.
+\end{equation*}
+If $M^2 < n^{-1}$, then we take $K = 1$, and from~\eqref{pf:laplacian_eigenmaps_estimation_fo_2} we get
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 \leq \frac{C}{n\delta}.
+\end{equation*}
+Finally if $M > n^{1/d}$, we take $K = n$. In this case, we note that $\wh{f}(X_i) = Y_i$ for all $i = 1,\ldots,n$, and it immediately follows that
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 = \frac{1}{n}\sum_{i = 1}^{n} w_i^2 \leq 5,
+\end{equation*}
+with probability at least $1 - \exp(-n)$. Combining these three separate cases yields the conclusion of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}.
+
+\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}.}
+Follows identically to the proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~\ref{prop:graph_seminorm_ho} rather than Proposition~\ref{prop:graph_seminorm_fo} and Assumption~\ref{asmp:parameters_estimation_ho} rather than Assumption~\ref{asmp:parameters_estimation_fo}.
+
+\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_manifold}.}
+Follows identically to the proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~\ref{prop:graph_seminorm_manifold} rather than Proposition~\ref{prop:graph_seminorm_fo}, Proposition~\textcolor{red}{(?)} rather than Proposition~\ref{prop:graph_eigenvalue}, and Assumption~\ref{asmp:parameters_estimation_manifold} rather than Assumption~\ref{asmp:parameters_testing_fo}.
+
+\subsection{Testing Results}
+
+\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}.}
+We have already upper bounded the Type I error of $\varphi$ in Lemma~\ref{lem:fixed_graph_testing}, and it remains to upper bound the Type II error.
+To do so, we condition on the event that the design points $X_1,\ldots,X_n$ satisfy,
+\begin{equation}
+\label{pf:laplacian_eigenmaps_testing_fo_1}
+\dotp{L_{n,\varepsilon}f_0}{f_0}_n \leq \frac{C}{\delta}M^2,\quad \textrm{and} \quad \lambda_k \geq \min\{\lambda_k(\Delta_P), \varepsilon^{-2}\}~~\textrm{for all $2 \leq k \leq n$,}
+\end{equation}
+as well as that
+\begin{equation}
+\label{pf:laplacian_eigenmaps_testing_fo_2}
+\|f_0\|_n^2 \geq \frac{1}{2}\|f_0\|_P^2.
+\end{equation}
+Note that by Propositions~\ref{prop:graph_seminorm_fo} and~\ref{prop:graph_eigenvalue}, both statements in~\eqref{pf:laplacian_eigenmaps_testing_fo_1} are satisfied with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\}$.
+Additionally, by Proposition~\ref{prop:empirical_norm_sobolev} and the assumption in~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} that $\|f_0\|_P^2 \geq CM^2/(bn^{2/d})$, the one-sided inequality~\eqref{pf:laplacian_eigenmaps_testing_fo_2} follows with probability at least $1 - \exp\{-(cn \wedge 1/b)\}$.
+Setting $\delta = b/3$ and taking $n \geq N$ to be sufficiently large, the bottom line is that both~\eqref{pf:laplacian_eigenmaps_testing_fo_1} and~\eqref{pf:laplacian_eigenmaps_testing_fo_2} are together satisfied with probability at least $1 - b/2$.
+
+Now, to complete the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}, we would like to invoke Lemma~\ref{lem:fixed_graph_testing}, and conclude that conditional on $X_1,\ldots,X_n$ satisfying~\eqref{pf:laplacian_eigenmaps_testing_fo_1} and~\eqref{pf:laplacian_eigenmaps_testing_fo_2}, our test $\varphi$ will equal $1$ with probability at least $1 - b/2$.
+To use Lemma~\ref{lem:fixed_graph_testing}, we will need to establish that~\eqref{eqn:fixed_graph_testing_critical_radius} is satisfied, which we now show.
+
+On the one hand, we have that the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius} is upper bounded, 
+\begin{align*}
+\frac{\dotp{L_{n,\varepsilon}f_0}{f_0}_n}{\lambda_{K + 1}} + \frac{\sqrt{2K}}{n}\biggl[2\sqrt{\frac{1}{a}} + \sqrt{\frac{2}{b}} + \frac{32}{bn}\biggr] & \leq C\biggl(\frac{M^2}{b \min\{\lambda_{K+1}(\Delta_P), \varepsilon^{-2}\}} + \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr) \\
+& \leq C\biggl(\frac{M^2}{b}K^{-2/d} + \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr)
+\end{align*}
+with the second inequality following by the assumption $\varepsilon \leq K^{-1/d}$ and Weyl's Law.
+On the other hand, we have that $\|f_0\|_n^2 \geq \|f_0\|_P^2/2$.
+Consequently, to prove Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}, it remains only to verify that
+\begin{equation}
+\label{pf:laplacian_eigenmaps_testing_fo_3}
+\|f_0\|_P^2 \geq C\biggl(\frac{M^2}{b}K^{-2/d} + \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr).
+\end{equation}
+As in the estimation case, we can further upper bound the right hand side of~\eqref{pf:laplacian_eigenmaps_testing_fo_3}, depending on the value of $K$ chosen in~\ref{asmp:parameters_testing_fo}.
+The classical case is $K = (M^2n)^{d/(2 + d)}$, in which case~\eqref{pf:laplacian_eigenmaps_testing_fo_3} is satisfied as long as
+\begin{equation*}
+\|f_0\|_P^2 \geq CM^2(M^2n)^{-4/(4 + d)}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]
+\end{equation*}
+If $M^2 < n^{-1}$, then we take $K = 1$, and~\eqref{pf:laplacian_eigenmaps_testing_fo_3} is satisfied whenever
+\begin{equation*}
+\|f_0\|_P^2 \geq \frac{C}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr].
+\end{equation*}
+Finally if $M > n^{1/d}$, we take $K = n$, and~\eqref{pf:laplacian_eigenmaps_testing_fo_3} is satisfied if
+\begin{equation*}
+\|f_0\|_P^2 \geq C\biggl(\frac{M^2}{n^{2/d}b} + n^{-1/2}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr).
+\end{equation*}
+We conclude by observing that~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} implies each of these three inequalities, and thus implies~\eqref{pf:laplacian_eigenmaps_testing_fo_3}.
+
+\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_ho}.}
+Follows identically to the proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~\ref{prop:graph_seminorm_ho} rather than Proposition~\ref{prop:graph_seminorm_fo} and Assumption~\ref{asmp:parameters_testing_ho} rather than Assumption~\ref{asmp:parameters_testing_fo}.
+
+\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_manifold}.}
+Follows identically to the proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~\ref{prop:graph_seminorm_manifold} rather than Proposition~\ref{prop:graph_seminorm_fo}, Proposition~\ref{prop:graph_eigenvalue_manifold} rather than Proposition~\ref{prop:graph_eigenvalue}, Proposition~\ref{prop:empirical_norm_sobolev_manifold} rather than Proposition~\ref{prop:empirical_norm_sobolev}, and Assumption~\ref{asmp:parameters_testing_manifold} rather than Assumption~\ref{asmp:parameters_testing_fo}.
+
+\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}.}
+Note that our choices of $K$ and $\varepsilon$ ensure  that~\eqref{pf:laplacian_eigenmaps_testing_fo_1} (with $L_{n,\varepsilon}^s$ replacing $L_{n,\varepsilon}$) and~\eqref{pf:laplacian_eigenmaps_testing_fo_2} are satisfied with probability at least $1 - b/2$.
+Proceeding as in the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}, we upper bound the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius},
+\begin{align*}
+\frac{\dotp{L_{n,\varepsilon}f_0}{f_0}_n}{\lambda_{K + 1}} + \frac{\sqrt{2K}}{n}\biggl[2\sqrt{\frac{1}{a}} + \sqrt{\frac{2}{b}} + \frac{32}{bn}\biggr] & \leq C\biggl(\frac{M^2}{b \min\{\lambda_{K+1}(\Delta_P), \varepsilon^{-2}\}} + \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr) \\
+& \leq C\biggl(\frac{M^2}{b}\varepsilon^2 + \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr).
+\end{align*}
+Unlike in the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}, we note that in this case $\varepsilon^2 \leq C\lambda_K(\Delta_P)$ rather than vice versa.
+From here, proceeding as in the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} gives the claimed result.
 
@@
 \begin{equation*}
-\lim_{n \to \infty} \sup_{h'} \frac{\|T_{h_n,n}\wc{f} - f_0\|_P^2}{\|T_{h',n}Y - f_0\|_P^2 } = 0,\quad \textrm{with probability $1$.}
+\lim_{n \to \infty} \sup_{h'} \frac{\Ebb \|T_{h_n,n}\wh{f} - f_0\|_P^2}{\Ebb \|T_{h',n}Y - f_0\|_P^2 } = 0.
 \end{equation*}
@@
 
-\subsection{Proof of Lemma~\ref{lem:kernel_smoothing_bias}}
+\textcolor{red}{(TODO)}
+
+\subsection{Kernel smoothing bias}
 \label{subsec:pf_kernel_smoothing_bias}
+
+\begin{lemma}
+	\label{lem:kernel_smoothing_bias}
+	Suppose Model~\ref{def:model_flat_euclidean}, and that the smoothing operator $T_{h,P}$ is computed with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel}.
+	\begin{itemize}
+		\item If $f_0 \in H^1(\mc{X})$, it follows that
+		\begin{equation*}
+		\|T_{h,P}f_0 - f_0\|_P^2 \leq C h^{2} |f|_{H^1(\mc{X})}^2.
+		\end{equation*}
+		\item If $f_0 \in H_0^{s}(P)$, $p \in C^{s}(\mc{X})$, and $\psi$ additionally satisfies~\ref{asmp:order_s_kernel}, it follows that
+		\begin{equation*}
+		\|T_{h,P}f_0 - f_0\|_P^2 \leq C h^{2s} |f|_{H^s(\mc{X})}^2.
+		\end{equation*}
+	\end{itemize}
+\end{lemma}
 
@@
 
-\textcolor{red}{(TODO)}
+In this section, we give a sequence of densities $p^{(n)}$ and regression functions $f_0^{(n)}$ for which the $L^2(P)$ error of the two-stage estimator $T_{h,n}\wh{f}$ is stochastically dominated by the $L^2(P)$ error of directly kernel smoothing.
+We begin by precisely setting up the problem and reviewing the two estimators we will compare.
+Then we give and prove our formal claim.
+
+\paragraph{Setup.}
+As usual, we assume the design points $(X_1,\ldots,X_n)$ are sampled independently from a distribution $P$, and the responses
+\begin{equation*}
+Y_i = f_0(X_i) + \varepsilon_i,
+\end{equation*}
+where the noise terms $\varepsilon_i \sim N(0,1)$ are sampled independently, and are also independent of the design points.
+
+Unlike in our main text, we shall assume a specific (parametric) form for $p = p^{(n)}$ and $f_0 = f_0^{(n)}$, that changes as a function of $n$. More specifically, using the notation
+\begin{equation*}
+r := \frac{(\log n)^2}{n}, \quad Q_1 := [0,1/2 - r], \quad Q_2 := [1/2 + r,1],
+\end{equation*}
+we take the domain $\mc{X}^{(n)} = Q_1 \cup Q_2$, and 
+\begin{equation}
+\label{def:model_cluster_assumption}
+p^{(n)}(x) := \frac{1}{1 - 2r}\1\bigl\{x \in Q_1 \cup Q_2\bigr\}, \quad f_0^{(n)}(x) := \theta \cdot \Bigl(\1\bigl\{x \in Q_1\bigr\} - \1\bigl\{x \in Q_2\bigr\}\Bigr),
+\end{equation}
+where $\theta \in \Reals$.
+Note that as $n \to \infty$, the density $p^{(n)}$ is converging (in, say, $L^2([0,1])$ norm) to uniform distribution on $[0,1]$, and the regression function $f_0^{(n)}$ to a step function.
+
+Our goal is to estimate $f_0$ using an estimator $f$, in such a way that the resulting $L^2(P)$ error $\|f - f_0\|_P^2$ is small.
+We shall consider two estimators: Laplacian eigenmaps + extension by kernel smoothing, and direct kernel smoothing of the responses.
+In the notation of our main text, these are respectively $T_{n,h}\wh{f}$ and $T_{n,h'}{\bf Y}$.
+For simplicity, in both cases we shall consider only the boxcar kernel,
+\begin{equation}
+\label{asmp:boxcar_kernel}
+\eta(z) = \psi(z) = \1\{z \leq 1\}.
+\end{equation}
+
+\paragraph{Laplacian eigenmaps + kernel smoothing beats direct kernel smoothing.}
+Now we are in a position to state our result. Let $P^{(n)}$ be the distribution with density $p^{(n)}$, i.e. the uniform probability distribution on $Q_1 \cup Q_2$. 
+\begin{proposition}
+	\label{prop:eigenmaps_beats_kernel_smoothing}
+	Suppose $(X_1,Y_1),\ldots(X_n,Y_n)$ are sampled according to~\eqref{def:model_cluster_assumption}. 
 	\begin{itemize}
-	\item \textcolor{red}{Mention the relationship between the estimation problem under consideration and classification with a margin (Tsybakov noise) condition.}
+		\item Compute the Laplacian eigenmaps estimator $\wh{f}$ using a kernel $\eta$ which satisfies~\eqref{asmp:boxcar_kernel}, number of eigenvectors $K = 2$, and radius $\varepsilon = ((\log n)^2/n)/2$.
+Compute the kernel smoothed extension $T_{n,h}\wh{f}$ using a kernel $\psi$ which satisfies~\eqref{asmp:boxcar_kernel}, and bandwidth $h = ((\log n)^2/n)/2$.
+Then,
+		\begin{equation}
+		\label{eqn:eigenmaps_beats_kernel_smoothing_1}
+		\Ebb \|T_{n,h}\wh{f} - f_0^{(n)}\|_{L^2(P^{(n)})}^2 \leq \frac{8 (\log n)^2 + 6 (\log n)^2 \theta^2}{n}.
+		\end{equation}
+		\item Compute the kernel smoothing estimator $T_{n,h}{\bf Y}$, using a kernel $\psi$ which satisfies~\eqref{asmp:boxcar_kernel}. Then
+		\begin{equation}
+		\label{eqn:eigenmaps_beats_kernel_smoothing_2}
+		\inf_{h' > 0} \Ebb\Bigl[\|T_{n,h'}{\bf Y} - f_0^{(n)}\|_{L^2(P^{(n)})}^2\Bigr] \geq \min\biggl\{ \frac{1}{16(\log n)^2}, \frac{\theta}{\sqrt{32 n}} \biggr\}.
+		\end{equation}
 	\end{itemize}
+\end{proposition}
+The statements~\eqref{eqn:eigenmaps_beats_kernel_smoothing_1} and~\eqref{eqn:eigenmaps_beats_kernel_smoothing_2} are finite sample and hold for any $\theta$.
+Taking a sequence $(\theta_n)_{n \in \mathbb{N}}$ such that $\theta = o(\sqrt{n})$ and $\theta = \omega((\log n)^2/\sqrt{n})$ implies that
+\begin{equation*}
+\lim_{n \to \infty} \sup_{h'}  \frac{\Ebb \|T_{n,h}\wh{f} - f_0^{(n)}\|_{P^{(n)}}^2}{	\Ebb \|T_{n,h'}{\bf Y} - f_0^{(n)}\|_{P^{(n)}}^2 } = 0.
+\end{equation*}
+
+\paragraph{Proof of Proposition~\ref{prop:eigenmaps_beats_kernel_smoothing}.}
+First we prove~\eqref{eqn:eigenmaps_beats_kernel_smoothing_1}, then~\eqref{eqn:eigenmaps_beats_kernel_smoothing_2}.
+
+\underline{\emph{Proof of~\eqref{eqn:eigenmaps_beats_kernel_smoothing_1}}.}
+We begin by showing that with high probability the eigenvectors $v_1,v_2$ respect the cluster structure of $p$.
+Formally, letting $u_1 = (\1\{X_i \in Q_1\})_{i \in [n]}$, and likewise $u_2 = (\1\{X_i \in Q_2\})_{i \in [n]}$, we establish that
+\begin{equation}
+\label{pf:eigenmaps_beats_kernel_smoothing_1}
+\Pbb\Bigl(\mathrm{span}\{v_1,v_2\} = \mathrm{span}\{u_1,u_2\}\Bigr) \geq 1 - \frac{1}{n}.
+\end{equation}
+Note that~\eqref{pf:eigenmaps_beats_kernel_smoothing_1} is equivalent to the statement that the neighborhood graph $G_{n,\varepsilon}$ consists of exactly two connected components: one consisting of all design points $X_i \in Q_1$, and the other consisting of all design points $X_i \in Q_2$.
+Of course, since $h < r$ and $\eta$ is compactly supported on $[0,1]$, it follows that no $X_i \in Q_1$ and $X_j \in Q_2$ can be connected.
+On the other hand, using an elementary concentration argument (stated in Lemma~\ref{lem:balls_in_bins}) and the triangle inequality, we deduce that with probability at least $1 - 1/n$ there exists a path in $G_{n,\varepsilon}$ between each $X_i, X_j \in Q_1$, and likewise between each $X_i,X_j \in Q_2$.
+This establishes~\eqref{pf:eigenmaps_beats_kernel_smoothing_1}.
+
+Now we condition on the ``good'' event $\mc{E}$ that the design points $X_1,\ldots,X_n$ satisfy~\eqref{eqn:balls_in_bins}, and therefore that $\mathrm{span}\{v_1,v_2\} = \mathrm{span}\{u_1,u_2\}$.
+Consider the empirical mean $\wb{Y}_Q := \frac{1}{\sharp\{Q \cup {\bf X}\}} \sum_{i: X_i \in Q} Y_i$.
+Since $\mathrm{span}\{v_1,v_2\} = \mathrm{span}\{u_1,u_2\}$, the estimator $\wh{f}$ will be piecewise constant on $Q_1$ and $Q_2$, and in fact we have that
+\begin{equation*}
+\wh{f} = \wb{Y}_{Q_1}u_1 + \wb{Y}_{Q_2}u_2.
+\end{equation*}
+Moreover, by~\eqref{eqn:balls_in_bins} every $x \in Q_1$ is within distance $h$ of at least one $X_i \in {\bf X} \cup Q_1$ and likewise every $x \in Q_2$ is within distance $h$ of at least one $X_j \in {\bf X} \cup Q_2$.
+Consequently, the kernel smoothening of $\wh{f}$ also exhibits a piecewise constant structure,
+\begin{equation*}
+T_{n,h}\wh{f}(x) = \wb{Y}_{Q_1}\1\{x \in Q_1\} + \wb{Y}_{Q_2}\1\{x \in Q_2\},\quad \textrm{for all}~~x \in Q_1 \cup Q_2.
+\end{equation*}
+Letting
+\begin{equation*}
+\|T_{n,h}\wh{f} - f_0^{(n)}\|_{L^2(P^{(n)})}^2 = \frac{1}{2}\Bigl((\wb{Y}_{Q_1} - \theta)^2 + (\wb{Y}_{Q_2} + \theta)^2\Bigr),
+\end{equation*}
+it follows from Lemma~\ref{lem:chi_square_bound} that conditional on $\mc{E}$,
+\begin{equation}
+\label{pf:eigenmaps_beats_kernel_smoothing_0}
+\|T_{n,h}\wh{f} - f_0^{(n)}\|_{L^2(P^{(n)})}^2 \leq \frac{4(\log n)^2}{n}.
+\end{equation}
+
+Now we derive a crude upper bound on $\|T_{n,h}\wh{f} - f_0^{(n)}\|_{L^2(P^{(n)})}$ that will suffice to control the error conditional on $\mc{E}^c$. Note that if $d_{n,h}(x) = 0$ then
+$T_{n,h}\wh{f}(x) = 0$, and $\bigl(T_{n,h}\wh{f}(x) - f_0(x)\bigr)^2 = [f_0(x)]^2$. On the other hand, if $d_{n,h}(x) \geq 1$, then by Jensen's inequality it follows that
+\begin{align*}
+\bigl(T_{n,h}\wh{f}(x) - f_0(x)\bigr)^2 & \leq \biggl(\frac{1}{d_{n,h}(x)} \sum_{i = 1}^{n} (\wh{f}_i - f_0(x)) \cdot \1\{\|X_i - x\| \leq h \}\biggr)^2 \\
+& \leq \frac{1}{d_{n,h}(x)} \sum_{i = 1}^{n} (\wh{f}_i - f_0(x))^2 \cdot \1\{\|X_i - x\| \leq h \} \\
+& \leq 2[f_0(x)]^2 + \frac{2}{d_{n,h}(x)} \sum_{i = 1}^{n} (\wh{f}_i)^2 \cdot \1\{\|X_i - x\| \leq h\},
+\end{align*}
+and integrating over all $x \in \mc{X}^{(n)}$ gives
+\begin{align*}
+\|T_{n,h}\wh{f} - f_0\|_{\Leb^2(P^{(n)})}^2 & \leq 2 \|f_0\|_{\Leb^2(P^{(n)})}^2 + \sum_{i = 1}^{n} (\wh{f}_i)^2 \biggl(\int \frac{\1\{\|X_i - x\| \leq h \}}{d_{n,h}(x)} \,dP(x)\biggr) \\
+& \leq 2 \|f_0\|_{\Leb^2(P^{(n)})}^2 + 2h\sum_{i = 1}^{n} (\wh{f}_i)^2 \\
+& = 2 \|f_0\|_{\Leb^2(P^{(n)})}^2 + 2 (\log n)^2 \|\wh{f}\|_n^2.
+\end{align*}
+We can further upper bound the empirical norm of $\|\wh{f}\|_n^2$ using the Cauchy-Schwarz inequality:
+\begin{equation*}
+\|\wh{f}\|_n^2 \leq \frac{2}{n}\sum_{i = 1}^{n}\dotp{Y}{v_1}_n^2v_{1,i}^2 + \dotp{{\bf Y}}{v_2}_n^2v_{2,i}^2 \leq 2\bigl(\dotp{{\bf Y}}{v_1}_n^2 + \dotp{{\bf Y}}{v_2}_n^2\bigr) \leq 4 \|{\bf Y}\|_n^2.
+\end{equation*}
+Noting that $\Ebb[\|{\bf Y}\|_n^2|{\bf X}] = \|f_0\|_n^2 + 1/n$, and that $\max_x |f_0(x)| = \theta$, we conclude that
+\begin{equation*}
+\Ebb\Bigl[\|T_{n,h}\wh{f} - f_0\|_{\Leb^2(P^{(n)})}^2 \cdot \1\{\mc{E}^c\}\Bigr] \leq \Ebb\Bigl[\Bigl(2 \|f_0\|_{\Leb^2(P^{(n)})}^2 + 4 (\log n)^2 \|{\bf Y}\|_n^2\Bigr) \cdot \1\{\mc{E}^c\} \Bigr] \leq \frac{2 \theta^2}{n} + \frac{4 (\log n)^2(\theta^2 + 1)}{n}.
+\end{equation*}
+Combining this with~\eqref{pf:eigenmaps_beats_kernel_smoothing_0} implies~\eqref{eqn:eigenmaps_beats_kernel_smoothing_1}.
+
+\underline{\emph{Proof of~\eqref{eqn:eigenmaps_beats_kernel_smoothing_2}}.}
+Fix $x \in \mc{X}^{(n)}$. A standard argument using the law of iterated expectation implies the following lower bound on the pointwise risk in terms of squared-bias and variance-like quantities,
+\begin{equation*}
+\Ebb\Bigl[\Bigl(T_{n,h'}{\bf Y}(x) - f_0(x)\Bigr)^2\Bigr] \geq \Ebb\biggl[\Bigl(f_0(X) - f_0(x)\Bigr)^2|X \in B(x,h')\biggr] + \Ebb\biggl[\frac{1}{d_{n,h'}(x)}\biggr].
+\end{equation*}
+The variance term can be lower bounded quite simply for any $x \in \mc{X}^{(n)}$; noting that $\sup_{x} p^{(n)}(x) < 2$ and $\nu(B(x,h') \cap \mc{X}^{(n)}) \leq 2h'$, it follows by Jensen's inequality that
+\begin{equation*}
+\Ebb\biggl[\frac{1}{d_{n,h'}(x)}\biggr] \geq \frac{1}{\Ebb[d_{n,h'}(x)]} \geq \frac{1}{4nh'}.
+\end{equation*}
+On the other hand the squared bias term is quite large for $x$ close to $1/2$. Precisely, if $h' \geq 4r$ then a simple calculation implies
+\begin{equation*}
+\Ebb[(f_0(X) - f_0(x))^2|X \in B(x,h')] \geq \frac{\theta^2}{8} \quad \textrm{for all}~x \in [(1 - h'/2)_{+}, 1/2 - r].
+\end{equation*}
+Combining these lower bounds on variance and squared bias terms and integrating over $x' \in \mc{X}$, we deduce the following: if $h' \leq 4r$, then
+\begin{equation*}
+\Ebb\Bigl[\|T_{n,h'}{\bf Y} - f_0^{(n)}\|_{L^2(P^{(n)})}^2\Bigr] \geq \frac{1}{16(\log n)^2},
+\end{equation*}
+whereas if $h' > 4r$  then
+\begin{equation*}\Ebb\Bigl[\|T_{n,h'}{\bf Y} - f_0^{(n)}\|_{L^2(P^{(n)})}^2\Bigr] \geq \frac{1}{4nh'} + \frac{\theta^2h'}{32}.
+\end{equation*}
+In the latter case, setting the derivative equal to $0$ shows that the right hand side is always at least $\theta/\sqrt{32 n}$, and taking the minimum over the two cases then yields~\eqref{eqn:eigenmaps_beats_kernel_smoothing_2}.
+
+\paragraph{A Useful Pair of Lemmas.}
+Let $m = 4n/\log^2n$. Furthermore, for $i = 0,\ldots,m - 1$, let
+\begin{equation*}
+Q_{i1} = [i/m,(i + 1)/m] \cdot (1/2 - r), \quad Q_{i2} = 1/2 + [i/m,(i + 1)/m] \cdot (1/2 - r).
+\end{equation*}
+\begin{lemma}
+	\label{lem:balls_in_bins}
+	Suppose $(X_1,Y_1),\ldots(X_n,Y_n)$ are sampled according to~\eqref{def:model_cluster_assumption}. For any $\log(n) > 16$, we have that 
+	\begin{equation}
+	\label{eqn:balls_in_bins}
+	\Pbb\Bigl(\sharp\{Q_{ij} \cup {\bf X}\} > 0 ~~\textrm{for all $i = 1,\ldots,m - 1$ and $j = 1,2$} \Bigr) \geq 1 - \frac{1}{n}.
+	\end{equation}
+\end{lemma}
+Let $\varepsilon = h = \log^2n/2n$.
+Note that by construction,~\eqref{eqn:balls_in_bins} implies that any points $x$ and $x'$ in adjacent intervals $Q_{ij}$ and $Q_{i'j}$ must be connected in $G_{n,\varepsilon}$.
+It also implies that $d_{n,h}(x) > 0$ for every $x \in Q_1 \cup Q_2$.
+
+The following Lemma is from~\cite{laurent00}.
+\begin{lemma}
+	\label{lem:chi_square_bound}
+	Let $\xi_1,\ldots,\xi_N$ be independent $N(0,1)$ random variables, and let $U := \sum_{k = 1}^{N} a_k(\xi_k^2 - 1)$.  Then for any $t > 0$,
+	\begin{equation*}
+	\Pbb\Bigl[U \geq 2 \norm{a}_2 \sqrt{t} + 2 \norm{a}_{\infty}t\Bigr] \leq \exp(-t).
+	\end{equation*}
+	In particular if $a_k = 1$ for each $k = 1,\ldots,N$, then
+	\begin{equation*}
+	\Pbb\Bigl[U\geq 2\sqrt{N t} + 2t\Bigr] \leq \exp(-t).
+	\end{equation*}
+\end{lemma}
+\paragraph{Proof (of Lemma~\ref{lem:balls_in_bins}).}
+For each $Q_{ij}$, we have that 
+\begin{equation*}
+\Pbb\bigl(\sharp\{Q_{ij} \cup {\bf X}\} = 0\bigr) = (1 - 1/(2m))^{n} \leq \exp\{-n/2m\}.
+\end{equation*}
+Thus by a union bound,
+\begin{equation*}
+\Pbb\Bigl(\sharp\{Q_{ij} \cup {\bf X}\} = 0 ~~\textrm{for any $i = 1,\ldots,m - 1$ and $j = 1,2$} \Bigr) \leq 2m\exp\{-n/2m\} = \frac{8n}{\log^2n}\exp\{-\log^2n/8\},
+\end{equation*}
+and the claim follows from some basic algebra.
 
@@
 Finally, we review notation and relevant facts regarding Taylor expansion.
+
+\subsection{Concentration Inequalities}
+
+The following Lemma is an immediate consequence of the one-sided Bernstein's inequality (14.23) in \cite{wainwright2019}.
+\begin{lemma}[One-sided Bernstein's inequality]
+	\label{lem:one_sided_bernstein}
+	Let $X, X_1,\ldots,X_n \sim P$, and $f$ satisfy $\Ebb[f^4(X)] < \infty$. Then with probability at least $1 - \exp\bigl(-n/8 \cdot \|f\|_P^4 /\Ebb[f^4(X)]\bigr)$,
+	\begin{equation*}
+	\|f\|_n^2 \geq \frac{1}{2}\|f\|_P^2.
+	\end{equation*}
+\end{lemma}
 
@@
 \textcolor{red}{(TODO): Check with Ryan and Siva that (i) they believe these facts, and (ii) these facts do not require additional justification.}
+
+\subsection{Upper bound on $L^2(P_n)$ risk}
+\label{subsec:upper_bound_l2pn_risk}
+
+\subsection{Facts about manifolds}
+\textcolor{red}{(TODO)}

### 2021-04-27 notes/misc/splines_graph_laplacians_connections.tex
Began a short note regarding connections between Laplacian smoothing splines and graph Laplacian smoothing.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=blue,
	citecolor=blue
}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{bm}
\usepackage{multirow}
\usepackage[font={small,it}]{caption}

\usepackage{natbib}
\renewcommand{\bibname}{References}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}
\DeclareMathAccent{\wc}{0}{mathx}{"71}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\RD}{\Reals^D}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}
\newcommand{\lambdavec}{\boldsymbol{\lambda}}
\newcommand{\bj}{{\bf j}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}
\newcommand{\Lap}{{\bf L}}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}
\newcommand{\Id}{I}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\SM}{\mathrm{SM}}
\newcommand{\OS}{\mathrm{FS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Note on: Connections between Smoothing Splines and Laplacian smoothing}
\author{Alden Green}
\date{\today}
\maketitle

Recall the definition of an \emph{order-$m$ smoothing spline}. For observations $(x_1,y_1),\ldots,(x_n,y_n)$, design points $x_i$ each belonging to a domain $\Omega \subset \Rd$, 
\begin{equation}
\label{eqn:smoothing_spline}
\wh{g} = \argmin_{g \in G} \frac{1}{n} \sum_{i = 1}^{n} (y_i - g(x_i))^2 + \lambda |g|_{m}^2;
\end{equation}
here we use $|g|_{m}$ to denote the $H^m(\Omega)$, and $G$ to be an appropriately defined function class. The standard choice is $G = H^m(\Omega)$, in which case we must also choose $m > d/2$ to make~\eqref{eqn:smoothing_spline} be well-defined.

The goal of this note is to review properties of order-$m$ smoothing splines, and see if/when graph Laplacian smoothing possesses analogous properties.

\section{Well-posedness of smoothing splines}
\textcolor{red}{(TODO)}

\section{Optimality of smoothing splines}
Suppose $x_1,\ldots,x_n$ are sampled independently from a \textcolor{red}{sufficiently regular} density $p$ on $\Omega$, and that
\begin{equation}
\label{model:regression}
y_i = f_0(x_i) + \varepsilon_i,
\end{equation}
where $\varepsilon_i \sim N(0,1)$ are i.i.d. Gaussian noise that are also independent of $x_i$. Assume that $f_0 \in H^s(\Omega)$. What can we say about mean squared error $\|\wh{g} - g\|_P^2$?

\textcolor{red}{(TODO)}:
\begin{itemize}
	\item Smoothing splines are optimal when $2m > s$. Is there a lower bound on $s$ relative to $m$ or $d$?
	\item When $m < s < 2s$, the optimality requires some boundary conditions on $f_0$. Detail these boundary conditions and explain intuition behind why they are required. Compare with the boundary conditions required for orthogonal series/non-parametric least squares estimators to be optimal.
 	\item Detail various techniques for establishing optimality. Refer to \textcolor{red}{(van der Geer, Nussbaum, Cox)}. Review pros and cons of each technique.
\end{itemize}

\section{Low-smoothness regime}
We now introduce a \emph{restricted smoothing spline}, in order to compute estimates for $m < d/2$. Suppose we take $\Phi_N \subseteq G$ to be a finite set of $L^2(P)$-orthogonal low-frequency functions; formally $\Phi_N = \{\phi_1,\ldots,\phi_N\}$, where each $\phi_k$ satisfies the Laplacian eigenfunction equation,
\begin{equation}
\label{eqn:laplacian_eigen}
\Delta_P \phi_k = \lambda_k \phi_k,\quad \frac{\partial}{\partial {\bf n}}\phi_k = 0.
\end{equation}
Let $F_N = \mathrm{span}\{\Phi_N\}$, and consider
\begin{equation}
\label{eqn:restricted_smoothing_spline}
\wh{g}_{\Phi} = \argmin_{g \in F_N} \frac{1}{n} \sum_{i = 1}^{n} (y_i - g(x_i))^2 + \lambda |g|_{m}^2.
\end{equation}
The restricted smoothing spline~\eqref{eqn:smoothing_spline} is well-posed, and its solution $\wh{g}_{\Phi}$ possesses a relatively simple analytic solution. It may be totally impossible to exactly compute because it is difficult to exactly solve~\eqref{eqn:laplacian_eigen}, but that is besides the point. Now we analyze its statistical properties. To do so, we reduce the analysis to that of a related estimator in the Normal means/Gaussian sequence model.

\textcolor{red}{(TODO)}
\begin{itemize}
	\item Refer to the results of \textcolor{blue}{(Birge and Massart)} for empirical risk minimization (ERM) over normed balls in H\"{o}lder spaces. Resolve why Birge and Massart find a phase transition at $m = 1, d = 2$, which is not at the $4m = d$ boundary; I suspect it is because Birge and Massart find the point at which the empirical process diverges.
\end{itemize}

\subsection{Gaussian sequence model}
In the Gaussian sequence model, for each $i \in \mathbb{N}$ one observes
\begin{equation}
\label{model:gaussian_sequence_model}
Y_i = \theta_i^{\ast} + \frac{1}{\sqrt{n}}\varepsilon_i,\quad \varepsilon_i \sim N(0,1).
\end{equation}
Supposes the sequence $\theta^{\ast}$ possesses some ``smoothness''. In particular, suppose $\theta^{\ast}$ belongs to the \emph{order-$s$ Sobolev ellipsoid} of radius $M$, $\Theta^s = \{\theta: |\theta|_s \leq M\}$, where for $s \in \mathbb{N}$ the Sobolev seminorm $|\theta|_s$ is defined by
\begin{equation*}
|\theta|_s^2 := \sum_{j = 1}^{\infty} \theta_j^2 j^{2s/d}.
\end{equation*}
Consider the \emph{sequence smoothing spline} estimator, defined for $N \in \mathbb{N}$ as
\begin{equation}
\label{eqn:sequence_smoothing_spline}
\wh{\theta}_{N} := \argmin_{\theta \in \ell^2(N)} |Y - \theta|_{0} + \lambda |\theta|_m^2, \quad \wh{\theta} := \lim_{N \to \infty} \wh{\theta}_{N}.
\end{equation}
Note that the sequence space estimator $\wh{\theta}$ is perfectly well-posed even though its analogous function space estimator is not. In fact, the sequence space estimator has the explicit solution
\begin{equation}
\wh{\theta}_{j} = Y_j[1 + \lambda j^{2m/d}]^{-1},
\end{equation}
showing that it is a shrinkage estimator.

\textcolor{red}{(TODO)}
\begin{itemize}
	\item Relate the sampling model~\eqref{model:gaussian_sequence_model} to~\eqref{model:regression}, and likewise relate the estimator~\eqref{eqn:sequence_smoothing_spline} to~\eqref{eqn:restricted_smoothing_spline}. Refer to \textcolor{blue}{(Tsybakov, Johnstone)}.
	\item Show that in the Gaussian sequence model, if $m > \max\{d/4,s/2\}$ then~\eqref{eqn:sequence_smoothing_spline} achieves optimal $\ell_2$ loss when $N = \infty$.
\end{itemize}
\subsection{``Failure'' of restricted smoothing splines in the very-low-smoothness regime}
On the other hand, whenever $m \leq d/4$ meaningful restrictions on the size of $N$ kick in.
\begin{lemma}
	\label{lem:failure_very_low_smoothness}
	Suppose we observe data $Y_1,Y_2,\ldots$ according to Model~\eqref{model:gaussian_sequence_model}. If we compute $\wh{\theta}$ and any $m \leq d/4$, then
	\begin{equation}
	\label{eqn:failure_very_low_smoothness}
	\Ebb[|\wh{\theta} - \theta^{\ast}|_{0}] = \infty.
	\end{equation}
\end{lemma}
The requirement $m \leq d/4$ is at first glance somewhat mysterious. It is nicely demystified by the following Lemma, which describes the expected ``smoothness'' properties of the estimator $\wh{\theta}$, in terms of the various Sobolev seminorms $|\cdot|_r$.
\begin{lemma}
	\label{lem:smoothness_of_sequence_smoothing_splines}
	Suppose we observe data $Y_1,Y_2,\ldots$ according to Model~\eqref{model:gaussian_sequence_model}, and we compute $\theta$ as in~\eqref{eqn:sequence_smoothing_spline}.
	\begin{itemize}
		\item If $4m > 2r + d$, then
		\begin{equation}
		\label{eqn:smoothness_of_sequence_smoothing_splines_1}
		\Ebb[|\wh{\theta}|_{r}^2] \leq |\theta^{\star}|_r^2 + O(\lambda^{-(2s + d)/(2m)}/n).
		\end{equation}
		\item If $4m \leq 2r + d$, then 
		\begin{equation}
		\label{eqn:smoothness_of_sequence_smoothing_splines_2}
		\Ebb[|\wh{\theta}|_{r}^2] = \infty.
		\end{equation}
	\end{itemize}
\end{lemma}
\begin{proof}
	Recall that $\wh{\theta}_i = Y_i \cdot [1 + \lambda i^{2m/d}]^{-1}$, and note that
	\begin{equation*}
	\Ebb[Y_i^2] = (\theta_i^{\ast})^2 + \frac{1}{n}.
	\end{equation*} 
	Therefore
	\begin{equation}
	\label{pf:smoothness_of_sequence_smoothing_splines_-1}
	\Ebb[|\wh{\theta}|_{r}^2] = \Ebb\biggl[\sum_{i = 1}^{\infty} \biggl(\frac{Y_i}{1 + \lambda i^{2m/d}}\biggr)^2 i^{2r/d}\biggr] = \sum_{i = 1}^{\infty} \frac{(\theta_i^{\ast})^2 + 1/n}{(1 + \lambda i^{2m/d})^2} \cdot i^{2r/d}
	\end{equation}
	Clearly,
	\begin{equation}
	\label{pf:smoothness_of_sequence_smoothing_splines_0}
	\sum_{i = 1}^{\infty} \frac{(\theta_i^{\ast})^2}{(1 + \lambda i^{2m/d})^2}  \cdot i^{2r/d} \leq \sum_{i = 1}^{\infty} (\theta_i^{\ast})^2 \cdot i^{2r/d} = |\theta^{\ast}|_{r}^2.
	\end{equation}
	Moreover, when $4m > 2r + d$, then
	\begin{equation}
	\label{pf:smoothness_of_sequence_smoothing_splines_1}
	\frac{1}{n}\sum_{i = 1}^{\infty} \frac{i^{2r/d}}{(1 + \lambda i^{2m/d})^2} \leq \frac{1}{n} \lambda^{-(2r + d)/(2m)}.
	\end{equation}
	On the other hand, when $4m \leq 2r + d$, then setting $i^{\ast}$ such that $(i^{\ast})^{2m/d}\lambda = 1$, we have
	\begin{equation}
	\label{pf:smoothness_of_sequence_smoothing_splines_2}
	\frac{1}{n}\sum_{i = 1}^{\infty} \frac{i^{2r/d}}{(1 + \lambda i^{2m/d})^2} \geq 	\frac{1}{2\lambda^2n}\sum_{i = i^{\ast}}^{\infty} i^{(2r - 4m)/d} = \infty.
	\end{equation}
	Plugging~\eqref{pf:smoothness_of_sequence_smoothing_splines_0}-\eqref{pf:smoothness_of_sequence_smoothing_splines_2} back into~\eqref{pf:smoothness_of_sequence_smoothing_splines_-1} establishes the claim.
\end{proof}
Lemma~\ref{lem:smoothness_of_sequence_smoothing_splines} has a particularly interesting interpretation when $r = 1$ or $r = 0$. 
\begin{itemize}
	\item When $r = 1$---corresponding to the first-order Sobolev norm--- Lemma~\ref{lem:smoothness_of_sequence_smoothing_splines} tells us that the sequence smoothing spline $\wh{\theta}$ has infinite $|\cdot|_1$ norm when $d \geq 2$. Remarkably, when $d = 2$ or $d = 3$, $\wh{\theta}$ continues to be optimal despite being an incredibly ``wiggly'' estimator of a ``smooth'' sequence. Note that this forecloses upon the possibility of analyzing $\wh{\theta}$ using uniform convergence over $\Theta^1(M)$, since $\wh{\theta}$ does not belong to this class for any finite $M$.
	\item When $r = 0$---corresponding to the zeroth-order Sobolev norm, or in other words the $\ell^2$-norm---Lemma~\ref{lem:smoothness_of_sequence_smoothing_splines} tells us that the sequence smoothing spline $\wh{\theta}$ has infinite $\Theta^0$ norm when $d \geq 4$. Clearly, since the regression function $f_0$ has finite $\Theta^0$ norm, and the estimator $\wh{f}$ has infinite $\Theta^0$ norm, it follows from the triangle inequality that $|\wh{f} - f|_0 = \infty$. 
\end{itemize}

\textcolor{red}{(TODO)}
\begin{itemize}
	\item Do an analogous calculation for the $\Theta^{1,1}$ case, with the estimator
	\begin{equation*}
	\wt{\theta}_{N} = \argmin_{\theta \in \ell^2(N)} |Y - \theta|_0^2 + \lambda |\theta|_{1,1}^2.
	\end{equation*}
	 Confirm your expectation: that when error measured in $|\cdot|_0$ norm and $d > 1$, this estimator is optimal when $N = n$, but not when $N = \infty$. This would give some evidence that the graph is saving the TV denoiser when $d > 1$.
	 \item Figure out other smoothness properties of $\wh{\theta}_N$, for instance ``H\"{o}lder'' smoothness.
	 \item Ponder what implication the lack-of-smoothness of sequence smoothing splines has for other problems (besides estimation with error measured in $|\cdot|_0$ norm).
\end{itemize}

\section{Relationship to Graph Laplacian methods}

\begin{itemize}
	\item Review why your analysis of Laplacian smoothing breaks down when $d > 4$, making reference to the previous section. If possible, prove a lower bound to this effect. Compare with the analogous situation for restricted smoothing splines, pointing out that the discretization imposed by the graph improves the situation slightly.
	\item Study whether order-$m$ Laplacian smoothing estimate is optimal over $H^1(\mc{X})$, as long as $m > d/4$. 
	\item Explain why your analysis of Laplacian smoothing fails at estimation over $H^s(\mc{X})$ for $s > 1$. If possible, prove a lower bound.
\end{itemize}

\end{document}

### 2021-04-28 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Completed the proof of Proposition 2 in Laplacian Eigenmaps paper.
Read through Laplacian Eigenmaps paper for errors. Ready to be merged.

@@
 \newcommand{\lambdavec}{\boldsymbol{\lambda}}
+\newcommand{\bj}{{\bf j}}
 
@@
 Geometric graphs encode both local information and global information about $P$ in an extremely general manner.
-For this reason they have been leveraged to conduct many different fundamental statistical tasks, such clustering, manifold learning, semi-supervised learning, classification, and regression \textcolor{red}{(TODO): references}.
+For this reason they have been leveraged to conduct many different fundamental statistical tasks, such as clustering, manifold learning, semi-supervised learning, classification, and regression \textcolor{red}{(TODO): references}.
 Substantial theoretical progress has been made regarding the consistency of \emph{graph-based learning}, that is, learning algorithms defined with respect to geometric graphs \textcolor{red}{(TODO): references}.
@@
 The specific graph-based method we study is \emph{Laplacian eigenmaps}, first introduced by \cite{belkin03a}, which projects the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto the span of eigenvectors of a graph Laplacian.
-We focus on the the unnormalized graph Laplacian $L$, which is a difference operator acting on functions $f: \{X_1,\ldots,X_n\} \to \Reals$ as follows,
+We focus on the the unnormalized graph Laplacian $L$, which is a difference operator acting on vectors $u \in \Reals^n$ as follows,
 \begin{equation}
 \label{eqn:graph_laplacian}
-Lf(X_i) = \sum_{j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)W_{ij}. 
-\end{equation}
-The Laplacian $L$ can be seen as a discretization of a weighted Laplace-Beltrami operator $\Delta_P$, defined when $P$ admits a density $p$ as
+(Lf)_i = \sum_{j = 1}^{n} (u_i - u_j)W_{ij}. 
+\end{equation}
+The graph Laplacian $L$ is a discrete approximation to a weighted continuum Laplacian operator $\Delta_P$, defined when $P$ admits a differentiable density $p$ as
 \begin{equation}
@@
 \end{equation}
-The eigenvectors of $L$ form an orthonormal basis of $L^2(P_n)$, and are approximations to the eigenfunctions of $\Delta_P$.
-Their corresponding eigenvalues are likewise estimates of eigenvalues of $\Delta_P$, and thus give a notion of smoothness to each eigenvector---roughly speaking, the smaller the eigenvalue, the smoother the corresponding eigenvector.
-
-Laplacian eigenmaps is thus a type \emph{spectral series} method.
-Spectral series methods, and more general orthogonal series methods, are some of most classical methods for non-parametric regression.
-Traditionally, orthogonal series methods are defined with respect to a fixed set of basis functions, orthogonalized in $L^2(Q)$ with respect to some pre-determined reference measure $Q$.
-Spectral series estimators take a particular orthogonal basis, the eigenfunctions of $\Delta_Q$.
-In contrast, eigenvectors of the graph Laplacian are data-dependent objects, and adapt to the geometry of the unknown design distribution $P$ in a rich manner.
-
-The theoretical properties of (classical) spectral series methods are various, and are in general extremely well-understood \textcolor{red}{(TODO): references}.
-In particular, they are minimax rate optimal for nonparametric regression over H\"{o}lder and Sobolev spaces.
-However, these methods suffer from some serious drawbacks.
-As a practical matter, even computing the eigenfunctions of $\Delta_Q$ is in general non-trivial.
-Moreover, the statistical optimality of such estimators also requires that the reference measure $Q$ be equal to $P$, whereas in this work we shall always assume $P$ is unknown.
-Of course there exist workarounds for these issues, for instance non-parametric least squares or \textcolor{red}{(...)}.
-We will comment more on these alternatives later; for now we merely emphasize that they constitute a fundamental change to the method under consideration.
-\textcolor{red}{Last sentence is weak, need to clarify, can maybe discuss with Ryan + Siva.}
-
-In contrast, Laplacian eigenmaps seeks to directly approximate a orthogonal series method that uses the eigenfunctions of $\Delta_P$, as we elaborate on momentarily.
-Importantly, it avoids the (arguably unrealistic) assumption that $P$ and the eigenfunctions of $\Delta_P$ be known, by using the graph Laplacian eigenvectors as a stand-in.
-On the other hand its theoretical statistical properties, and in particular its optimality as a method for regression, remain poorly understood.
+The eigenvectors of $L$ form an orthonormal basis of $\Reals^n$, and serve as estimates of eigenfunctions of $\Delta_P$.
+Their corresponding eigenvalues are likewise estimates of eigenvalues of $\Delta_P$, and give a notion of smoothness to each eigenvector---roughly speaking, the smaller the eigenvalue, the smoother the corresponding eigenvector.
+
+We can therefore view Laplacian eigenmaps as a twist on a very classical approach to nonparametric regression: \emph{spectral projection}, or more generally \emph{orthogonal series}, regression.
+Classically, in orthogonal series regression one fixes a reference measure $Q$, takes an ordered orthonormal basis $\psi_1,\psi_2,\ldots$ of $L^2(Q)$, computes empirical Fourier coefficients $\dotp{{\bf Y}}{\psi_k}_n$, and uses the first few terms in the resulting Fourier series to construct an estimator.
+Regression by spectral projection is a special case of this general setup, where one takes $\psi_k$ to be the $k$th eigenfunction of the continuum Laplacian operator $\Delta_Q$.
+In contrast to this classical approach---in which the reference measure $Q$ and resulting Laplacian $\Delta_Q$ are determined a priori---in Laplacian eigenmaps the eigenvectors of the graph Laplacian serve as the basis.
+These eigenvectors are data-dependent objects, and adapt to the geometry of the unknown design distribution $P$ in a rich manner.
+For instance, they respect the cluster structure of $P$, meaning that if the density $p$ has multiple connected components, the first few eigenvectors of $L$ will (with high probability) be piecewise constant over each such component \textcolor{blue}{(vonLuxburg 2009)}.
+On the other hand if $P$ is supported on some low-dimensional manifold, the graph Laplacian eigenvectors concentrate around the eigenfunctions of a manifold Laplace-Beltrami operator, and thus give a principled embedding of potentially high dimensional design points $X_i$ into a lower dimensional space \citep{belkin03a}.
+
+Thus Laplacian eigenmaps is a data-dependent alternative to classical spectral projection estimators.
+This data-dependency is appealing.
+Classical spectral projection estimators possess attractive theoretical properties for nonparametric regression---in particular, they are minimax rate-optimal for regression over H\"{o}lder and Sobolev spaces \cite{tsybakov08,johnstone2011,gine16}---but in practice suffer from some serious drawbacks.
+A basic difficulty is that finding the eigenfunctions of $\Delta_Q$ is in general non-trivial, so that it may not be possible to compute the estimator in the first place.
+Moreover, spectral projection estimators make sense (and are minimax optimal) only if the reference measure $Q$ is very close or equal to $P$, since otherwise the basis functions are orthogonalized with respect to the wrong measure.
+For these reasons, such estimators are typically proposed and studied under very restrictive conditions on the design points: for instance, that they are equally spaced fixed grid points, or that they are random but uniformly distributed on the unit cube $[0,1]^d$.
+There do exist fixes to the issues just raised, for instance using nonparametric least-squares.
+We will compare Laplacian eigenmaps to nonparametric least squares in more detail later, and for now merely point out that it fundamentally changes the estimator under consideration.
+\textcolor{red}{(For Ryan + Siva): I attempted to edit this but still don't like it.}
+
+In contrast, Laplacian eigenmaps directly approximates a spectral projection method: it projects the responses onto an orthogonal basis (eigenvectors of the graph Laplacian) that approximates the smooth eigenfunctions of $\Delta_P$.
+Laplacian eigenmaps is perfectly well-defined, and indeed straightforward to compute, when the design $P$ is unknown.
+This includes the situation where $P$ may be non-uniform, or even concentrated on a low-dimensional manifold.
+On the other hand, because graph Laplacian eigenvectors depend on the random design points $X_1,\ldots,X_n$ in \textcolor{red}{(non-trivial ways)}, it is substantially more difficult to analyze Laplacian eigenmaps than to analyze classical spectral projection methods.
+For this reason, the theoretical statistical properties of Laplacian eigenmaps, and in particular its optimality as a method for nonparametric regression with random design, remain poorly understood.
 
@@
 Broadly seeking, our answer is yes.
-We consider two different data models, one of which assumes a full-dimensional, flat Euclidean domain, and the other of which assumes that the distribution $P$ is supported on a manifold of small intrinsic dimension $m$.
-We show that when the regression function $f_0$ is smooth, in the Sobolev sense of having weak derivatives bounded in $\Leb^2$ norm, Laplacian eigenmaps methods are statistically minimax optimal, for both estimation and goodness-of-fit testing.
-Our statements holds for different relations between the dimension $d$ (or $m$) and number of derivatives $s$, depending on the problem (estimation or testing), and are summarized in Tables~\ref{tbl:estimation_rates} and~\ref{tbl:testing_rates}.
+We consider two different data models, one of which assumes the support $\mc{X}$ of the distribution $P$ is a full-dimensional, flat Euclidean domain, and the other of which assumes that $\mc{X}$ is a manifold of small intrinsic dimension $m$.
+We show that when the regression function $f_0$ is smooth, in the Sobolev sense of having weak derivatives bounded in $\Leb^2(\mc{X})$ norm, Laplacian eigenmaps methods are statistically minimax optimal, for both estimation and goodness-of-fit testing.
+Our statements hold for different relations between the dimension $d$ (or $m$) and number of derivatives $s$, depending on the problem (estimation or testing), and are summarized in Tables~\ref{tbl:estimation_rates} and~\ref{tbl:testing_rates}.
 \begin{table}
@@
 In each case, rates hold for all $d \in \mathbb{N}$ (under Model~\ref{def:model_flat_euclidean}), and for all $m \in \mathbb{N}, 1 < m < d$ (under Model~\ref{def:model_manifold}).
-Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps estimator is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $M \gtrsim n^{-1/2}$.}
+Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps estimator is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $n^{-1/2} \lesssim M \lesssim n^{1/d}$.}
 	\label{tbl:estimation_rates}
@@
 Rates when $d > 4s$ assume that $f_0 \in L^4(P,M)$.
-Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $M \gtrsim n^{-1/2}$.}
+Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $n^{-1/2} \lesssim M \lesssim n^{1/d}$.}
 	\label{tbl:testing_rates}
@@
 \paragraph{Related Work.}
-\textcolor{red}{(TODO)}: Flesh this out.
-\begin{itemize}
-	\item Summarize what is known on fixed graphs, drawing from our paper on Laplacian smoothing.
-	\item \textcolor{red}{von Luxburg, Belkin, Slepcev} show that for a fixed index $k$, the eigenvector $v_k$ converges to a eigenfunction $\psi_k$ of the Laplace-Beltrami operator $\Delta_P$.
-\textcolor{red}{Shi, Burago, Garcia Trillos, Calder, Cheng} build on this, giving finite sample bounds, rates of convergence, and making statements uniform over $1 \leq k \leq k_{\max}$.
-This makes it clear that Laplacian eigenmaps is approximating a classical spectral series method, and may be a good alternative when the latter cannot be implemented (e.g.
-because $P$ or $\mc{X}$ are unknown.) It is tempting to try to directly use these results along with the triangle inequality to deduce that if (i)  spectral series methods are optimal, and (ii) Laplacian eigenmaps is converging to a spectral series method, then (iii) Laplacian eigenmaps must be also optimal.
-Unfortunately, this won't work, both because we will need to take an increasing number of eigenvectors (increasing, that is, with the sample size $n$) to achieve an optimal estimator and test, and because the best known rates of convergence of $v_k \to \psi_k$ are slower than the minimax rates.
-	\item There has been limited analysis of the estimator $\wh{f}$.
-\cite{zhou2011} consider $\wh{f}$, but in the semi-supervised setting where the number of unlabeled points grows to infinity.
-In this case $\wh{f}$ reduces to a classical spectral series estimator, and there is no error incurred by approximating $\Delta_P$.
-\cite{lee2016} consider the \emph{diffusion maps} estimator---which uses the eigenvectors of a different normalization of the Laplacian $L$---in both the supervised and semi-supervised setups, but they derive suboptimal rates for the supervised problem.
+There is an incredible amount of work on the properties of classical spectral projection methods, which go far beyond their optimality for nonparametric regression.
+We will not attempt to summarize this work, and instead refer to \citet{wasserman2006,gyorfi2006,tsybakov08,johnstone2011,gine16} and the references therein.
+
+More recent work has considered regression on a fixed graph, where one treats the design points $x_1,\ldots,x_n$ as vertices in a fixed graph $G$, and carries out inference with respect to the function evaluations $(f_0(x_i))_{i = 1}^{n}$ By this point there exists a relatively mature theory describing this setting.
+Tight upper bounds have been established that certify the optimality of graph-based methods for both nonparametric estimation \citep{wang2016,hutter2016,sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}) and testing (e.g., \citet{sharpnack2010identifying,sharpnack2013b,sharpnack2013,sharpnack2015} over different ``function'' classes (in quotes because these classes really model the $n$-dimensional vector of evaluations) We call particular attention to~\citet{sadhanala16} and~\citet{sharpnack2015}, who analyze the Laplacian eigenmaps estimator and test statistic, respectively.
+This setting is quite general, because the graph need not be a geometric graph defined on a vertex set which belongs to Euclidean space.
+On the other hand, in many situations it may be somewhat to unnatural to assume that the design points are a priori fixed, and that the regression function $f_0$ exhibits ``smoothness'' over this fixed design.
+Instead, it may be more reasonable to adopt the \emph{random design} perspective that we work in, and assume that the regression function $f_0$ exhibits a more classical notion of smoothness.
+
+However, as already mentioned, when the design points are random so too are the graph Laplacian eigenvectors, and grasping their properties is in general non-trivial.
+For this reason, there has not been much analysis of random design nonparametric regression using Laplacian eigenmaps.
+\cite{zhou2011} consider the Laplacian eigenmaps estimator, but in the semi-supervised setting, where one additionally observes unlabeled design points $X_{n + 1},\ldots,X_{n + m}$.
+Their analysis assumed that for a fixed number of labeled samples $n$, the number of unlabeled samples $m$ grows to infinity.
+In this case the eigenvectors of the graph Laplacian converge to eigenfunctions of a continuum Laplacian, and the analysis of the resulting estimator is identical to that of a classical spectral projection estimator.
+\cite{lee2016} consider the \emph{diffusion maps} estimator---which uses the eigenvectors of a different normalization of the graph Laplacian $L$---in both the supervised and semi-supervised setups.
+In the supervised case, they show that the diffusion maps estimator converges to the regression function as the sample size $n \to \infty$, but at a suboptimal rate.
 As far as we know, there has been no analysis of the test statistic $\wh{T}$ in the random design framework which we study.
-	\item Alternatively, \cite{trillos2020} \textcolor{blue}{(Green 2021)} use the graph Laplacian operator to induce a penalty over functions $f: \{\bf X\} \to \Reals$.
-The \emph{Laplacian smoothing} estimator $\wh{f}_{\mathrm{LS}}$ is obtained by minimizing the sum of this penalty along with a data-fidelity term,
+
+There has been much more analysis of the convergence properties of eigenvectors of random graph Laplacians to their continuum limits.
+\citet{belkin07,vonluxburg2008,singer2017, garciatrillos18} show that eigenvalue-eigenvector pairs $(\lambda,v)$ of a graph Laplacian converge to eigenvalue-eigenfunction pairs $(\lambda,\psi)$ of a limiting differential or integral operator.
+\citet{burago2014, shi2015, trillos2019, calder2019, cheng2021} build on these works, by giving finite sample bounds, rates of convergence, and making statements uniform.
+These results justify our intuition that Laplacian eigenmaps, which projects the responses ${\bf Y}$ onto eigenvectors of a graph Laplacian, is approximating a classical spectral projection method, which projects ${\bf Y}$ onto eigenfunctions of the limiting differential operator $\Delta_P$.
+In fact, more formally these results imply that for a fixed number of eigenvectors $K > 0$, the Laplacian eigenmaps estimator converges to its classical counterpart as $n \to \infty$.
+By taking $K \to \infty$ sufficiently slowly with $n$, we can even conclude that Laplacian eigenmaps will be a consistent estimator of $f_0$.
+Unfortunately, the resulting rates of convergence implied by this approach are severely suboptimal, compared to the minimax rates.
+Our analysis showing that Laplacian eigenmaps achieves minimax optimal rates of convergence will use some of the results mentioned above, but will overall proceed by a very different route than the one just considered.
+
+Finally, we point out that there are other ways to use neighborhood graphs, and specifically graph Laplacians, for nonparametric regression.
+For instance, \cite{trillos2020} \textcolor{blue}{(Green 2021)} use the graph Laplacian to induce a penalty over functions $f: \{\bf X\} \to \Reals$.
+The \emph{Laplacian smoothing} estimator $\wh{f}_{\mathrm{LS}}$ is obtained by minimizing the sum of this penalty with a data-fidelity term,
 \begin{equation*}
@@
 \textcolor{blue}{(Green 2021)} show that the resulting estimator is minimax optimal, but only for $s = 1$ and $d \leq 4$.
-In contrast, the Laplacian eigenmaps estimator is optimal for all $s$ and $d$.
-\end{itemize}
+In contrast, we show in this work that the Laplacian eigenmaps estimator is optimal for all $s$ and $d$.
 
@@
 We now outline the structure of the rest of this paper.
-In Section~\ref{sec:setup_main_results}, we formally define the regression problem and estimator we consider, and give some background about minimax regression over Sobolev spaces.
+In Section~\ref{sec:setup_main_results}, we formally define the regression problem and estimator we consider.
+We also give some background about minimax regression over Sobolev spaces, and recall the classical spectral projection estimators which achieve minimax rates of convergence.
 In Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity}, we establish our main results regarding the optimality of Laplacian eigenmaps.
-The Laplacian eigenmaps estimator $\wh{f}$ is defined only at the design points $X_1,\ldots,X_n$.
-In Section~\ref{sec:out_of_sample}, we propose a method for out-of-sample extension of $\wh{f}$, and show that it has optimal out-of-sample error.
-In Section~\ref{sec:experiments} we examine the empirical behavior of Laplacian eigenmaps, and compare it to some natural competitors for non-parametric regression.
+Since thhe Laplacian eigenmaps estimator $\wh{f}$ is defined only at the design points $X_1,\ldots,X_n$, in Section~\ref{sec:out_of_sample} we propose a method for out-of-sample extension of $\wh{f}$, and show that it has optimal out-of-sample error.
+In Section~\ref{sec:experiments} we examine the empirical behavior of Laplacian eigenmaps, and compare it to some natural competitors for nonparametric regression.
 We conclude with some discussion in Section~\ref{sec:discussion}.
@@
 \paragraph{Notation.}
-
-\textcolor{red}{TODO}:
-
-\begin{itemize}
-	\item Define Lebesgue measure $\nu$. 
-	\item Define Euclidean norm $\|\cdot\|$.
-	\item Define $\Leb^q(\mc{X}), \Leb^2(P)$ and $\Leb^2(P_n)$. Define $C^k(\mc{X})$, and its norm. Define $C_0^k(\mc{X})$.
-	\item Define asymptotic notation: $\ll$, $\lesssim$ and $\asymp$.
-	\item Define min and max notation: $\vee$ and $\wedge$.
-\end{itemize}
+We frequently refer to various classical function classes.
+For a domain $\mc{X}$ with volume form $d\mu$, we let $L^2(\mc{X})$ denote the set of functions $f$ for which $\|f\|_{L^2(\mc{X})}^2 := \int f^2 \,d\mu  < \infty$, and equip $L^2(\mc{X})$ with the norm $\|\cdot\|_{L^2(\mc{X})}$.
+We define $\dotp{f}{g}_P := \int fg\,dP$, and let $L^2(P)$ contain those functions $f$ for which $\|f\|_P^2 := \dotp{f}{f}_P$ is finite.
+Finally, we let $L^2(P_n)$ consist of those ``functions'' $f: \set{X_1,\ldots,X_n} \to \Reals$ for which the empirical norm $\|f\|_{n}^2 := \frac{1}{n}\sum_{i = 1}^{n} \bigl(f(X_i)\bigr)^2 < \infty$.
+When there is no chance of confusion, we will sometimes associate functions in $L^2(P_n)$ with vectors in $\Reals^n$, and vice versa.
+We use $C^k(\mc{X})$ to refer to functions which are $k$ times continuously differentiable in $\mc{X}$, either for some integer $k \geq 1$ or for $k = \infty$.
+We let $C_c^{\infty}(\mc{X})$ represent those functions in $C^{\infty}(\mc{X})$ which are compactly contained in $\mc{X}$.
+We write $\partial f/\partial r_i$ for the partial derivative of $f$ in the $i$th standard coordinate of $\Rd$, and use the multi-index notation $D^{\alpha}f := \partial^{|\alpha|}f/\partial^{\alpha_1}x_1\ldots\partial^{\alpha_d}x_d$ for multi-indices $\alpha \in \Reals^m$.
+
+We write $\|\cdot\|$ for Euclidean, and $d_{\mc{X}}(x',x)$ for the geodesic distance between points $x$ and $x'$ on a manifold $\mc{X}$.
+Then for a given $\delta > 0$, $B(x,\delta)$ is the radius-$\delta$ ball with respect to Euclidean distance, whereas $B_{\mc{X}}(x,\delta)$ is the radius-$\delta$ ball with respect to geodesic distance.
+Letting $T_x(\mc{X})$ be the tangent space at a point $x \in \mc{X}$, we write $B_m(v,\delta) \subset T_x(\mc{X})$ for the radius-$\delta$ ball centered at $v \in T_x(\mc{X})$.
+
+For sequences $(a_n)$ and $(b_n)$, we use the asymptotic notation $a_n \lesssim b_n$ to mean that there exists a number $C$ such that $a_n \leq C b_n$ for all $n$.
+We write $a_n \asymp b_n$ when $a_n \lesssim b_n$ and $b_n \lesssim a_n$.
+On the other hand we write $a_n = o(b_n)$ when $\lim a_n/b_n = 0$, and likewise $a_n = \omega(b_n)$ when $\lim a_n/b_n = \infty$.
+Finally $a \vee b := \max\{a,b\}$ and $a \wedge b := \min\{a,b\}$.
 
@@
 
-In this section, we begin by giving a precise definition of our framework, and the Laplacian eigenmaps methods we study. We then review minimax rates for non-parametric regression over Sobolev spaces.
-
-\subsection{Non-parametric regression with random design}
+In this section, we begin by giving a precise definition of our framework, and the Laplacian eigenmaps methods we study.
+We then review minimax rates for nonparametric regression over Sobolev spaces.
+We pay special attention to the classical spectral projection methods which achieve these rates, since such methods are closely connected to Laplacian eigenmaps.
+
+\subsection{Nonparametric regression with random design}
 \label{sec:regression_laplacian_eigenmaps}
 
-We will operate in the usual setting of non-parametric regression with random design, in which we observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$.
-The design points $X_1,\ldots,X_n$ are sampled from a distribution $P$ with support $\mc{X} \subseteq \Rd$, and the responses follow
+We will operate in the usual setting of nonparametric regression with random design, in which we observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$.
+The design points $X_1,\ldots,X_n$ are sampled from a distribution $P$ with support $\mc{X} \subseteq \Rd$, and the responses follow the signal plus noise model
 \begin{equation}
@@
 The support $\mc{X}$ of the design distribution $P$ is an open, connected, and bounded subset of $\Rd$, with Lipschitz boundary.
-The distribution $P$ admits a density $p$ with respect to the $d$-dimensional Lebesgue measure $\nu$, which is bounded away from $0$ and $\infty$,
+The distribution $P$ admits a Lipschitz density $p$ with respect to the $d$-dimensional Lebesgue measure $\nu$, which is bounded away from $0$ and $\infty$,
 	\begin{equation*}
@@
 	The support $\mc{X}$ of the design distribution $P$ is a closed, connected, smooth and boundaryless Riemannanian manifold embedded in $\Rd$, of intrinsic dimension $1 \leq m < d$.
-The injectivity radius of $\mc{X}$ is upper bounded by a constant $i_0$.
-The design distribution $P$ admits a density $p$ with respect to the volume form $d\mu$ induced by the Riemannian structure of $\mc{X}$, which is bounded away from $0$ and $\infty$,
+The injectivity radius of $\mc{X}$ is lower bounded by a positive constant $i_0 > 0$.
+The design distribution $P$ admits a Lipschitz density $p$ with respect to the volume form $d\mu$ induced by the Riemannian structure of $\mc{X}$, which is bounded away from $0$ and $\infty$,
 	\begin{equation*}
@@
 \end{definition}
+
+Finally, at various points we will have to assume that the density $p$ also displays different types of higher-order regularity, beyond Lipschitz continuity.
+When such assumptions are necessary, we always will state them explicitly.
 
@@
 We now formally define the estimator and test statistic we study.
-Both are derived from eigenvectors of the neighborhood graph Laplacian operator.
-For a positive, symmetric kernel $\eta: [0,\infty) \to [0,\infty)$, and a bandwidth parameter $\varepsilon > 0$, the \emph{neighborhood graph Laplacian} operator $L_{n,\varepsilon}$ is defined as
+Both are derived from eigenvectors of a graph Laplacian.
+For a positive, symmetric kernel $\eta: [0,\infty) \to [0,\infty)$, and a radius parameter $\varepsilon > 0$, let $G = ([n],W)$ be the neighborhood graph formed over the design points $\{X_1,\ldots,X_n\}$, with a weighted edge $W_{ij} = \eta(\|X_i - X_j\|/\varepsilon)$ between vertices $i$ and $j$.
+Then the
+\emph{neighborhood graph Laplacian} $L_{n,\varepsilon}: \Reals^n \to \Reals$ is defined by its action on vectors $u \in \Reals^n$ as
 \begin{equation}
 \label{eqn:neighborhood_graph_laplacian}
-L_{n,\varepsilon}u(x) = \frac{1}{n\varepsilon^{2 + \mathrm{dim}(\mc{X})}} \sum_{i = 1}^{n} \bigl(u(x) - u(X_j)\bigr) \eta\biggl(\frac{\|x - X_j\|}{\varepsilon}\biggr)
+\bigl(L_{n,\varepsilon}u\bigr)_i := \frac{1}{n\varepsilon^{2 + \mathrm{dim}(\mc{X})}} \sum_{j = 1}^{n} \bigl(u_i - u_j\bigr) \eta\biggl(\frac{\|X_i - X_j\|}{\varepsilon}\biggr).
 \end{equation}
@@
 The pre-factor $(n\varepsilon^{2 + \mathrm{dim}(\mc{X})})^{-1}$ is purely for convenience in taking limits as $n \to \infty, \varepsilon \to 0$).
-
-Although~\eqref{eqn:neighborhood_graph_laplacian} makes sense for any $x \in \mc{X}$ and $u: \mc{X} \to \Reals$, we will typically think of $x \in \{X_1,\ldots,X_n\}$ and $u = (u(X_1),\ldots,u(X_n)) \in \Leb^2(P_n)$), so that $L_{n,\varepsilon}$ is equivalently a matrix in $\Reals^{n \times n}$.
-Indeed, viewed as a matrix $L_{n,\varepsilon}$ is a symmetric and positive semi-definite, and can thus be diagonalized by its eigenpairs, solutions $(\lambda,v)$ to the eigenproblem
-\begin{equation*}
-\frac{1}{n}L_{n,\varepsilon}v = \lambda v, \quad \|v\|_n^2 = 1.
+Written in standard coordinates we have $(n\varepsilon^{\dim(\mc{X}) + 2}) \cdot L_{n,\varepsilon} = D - W$, where $D \in \Reals^{n \times n}$ is the diagonal degree matrix, $D_{ii} = \sum_{i = 1}^{n} W_{ij}$.
+
+The graph Laplacian is a positive semi-definite matrix, and admits the eigendecomposition $L_{n,\varepsilon} = \sum_{k = 1}^{n} \lambda_k v_k v_k^{\top}$, where for each $k = 1,\ldots,n$ the eigenvalue-eigenvector pair $(\lambda_k,v_k)$ satisfies
+\begin{equation*}
+\frac{1}{n}L_{n,\varepsilon}v_k = \lambda_k v_k, \quad \|v_k\|_n^2 = 1.
 \end{equation*}
@@
 \end{equation} 
-If $\wh{f}$ is a reasonable estimate of $f_0$, then the Laplacian eigenmaps test statistic
+Thus $\wh{f}$ is equivalently a vector in $\Reals^n$, or a function in $L^2(P_n)$. If $\wh{f}$ is a reasonable estimate of $f_0$, then the Laplacian eigenmaps test statistic
 \begin{equation}
@@
 Laplacian eigenmaps typically refers to an algorithm for embedding, which maps each design point $X_1,\ldots,X_n$ to $\Reals^K$ according to $X_i \mapsto (v_{1,i}, \ldots, v_{K,i})$.
-Viewing this mapping as a feature map, we can then interpret the estimator $\wh{f}$ as the least-squares solution to a linear regression problem with responses $Y_1,\ldots,Y_n$ and features $v_1,\ldots,v_K$.
-Often, the Laplacian eigenmaps embedding is viewed as a tool for dimensionality reduction, wherein it is implicitly assumed that $K \ll d$.
+Viewing this embedding as a feature map, we can then interpret the estimator $\wh{f}$ as the least-squares solution to a linear regression problem with responses $Y_1,\ldots,Y_n$ and features $v_1,\ldots,v_K$.
+Often, the Laplacian eigenmaps embedding is viewed as a tool for dimensionality reduction, wherein it is implicitly assumed that $K$ is much smaller than $d$.
 We will neither explicitly nor implicitly take $K < d$; indeed, the embedding perspective is not particularly illuminating in what follows, and we do not henceforth make reference to it.
@@
 
-% AG: Said elsewhere. 
-% On a different note, it is worth explicitly pointing out that $\wh{f}$ is defined only in-sample, that is, only at the design points $X_1,\ldots,X_n$.
-There is no straightforward or agreed upon way to extend $\wh{f}$ out-of-sample, and we address the problem of out-of-sample estimation in some detail later in Section~\ref{sec:out_of_sample}.
-%
-
 \subsection{Sobolev Classes}
 \label{sec:sobolev}
-Before we give an overview of our main results, we will first pause to review the definition of Sobolev classes.
-We divide our discussion into two cases---the flat Euclidean case, and the manifold case, corresponding to Models~\ref{def:model_flat_euclidean} and Model~\ref{def:model_manifold}.
+We now review the definition of Sobolev classes, dividing our discussion into two cases---the flat Euclidean case and the manifold case, corresponding respectively to Models~\ref{def:model_flat_euclidean} and Model~\ref{def:model_manifold}.
 
 \paragraph{Flat Euclidean case.}
-Take $\mc{X}$ to be an open, connected set with Lipschitz boundary, as in Model~\ref{def:model_flat_euclidean}.
+Take $\mc{X}$ to be a open, connected and bounded set with Lipschitz boundary, as in Model~\ref{def:model_flat_euclidean}.
 Recall that for a given multiindex $\alpha \in \mathbb{N}^d$, a function $f$ is \emph{$\alpha$-weakly differentiable} if there exists some $h \in L^1(\mc{X})$ such that
 \begin{equation*}
-\int_{\mc{X}} h g = (-1)^{|\alpha|} \int_{\mc{X}} f D^{\alpha}g, \quad \textrm{for every $g \in C_0^{\infty}(\mc{X})$.}
-\end{equation*}
-If such a function $h$ exists, it is the $\alpha$th weak partial derivative of $f$, and denoted by $D^{\alpha}f = h$.
-If $f$ is $\alpha$-weakly differentiable for all $|\alpha| = j$, then $d^jf := \sum_{|\alpha| = j}D^{\alpha}f$ is the sum of all weak partial derivatives of order $j$.
-
-For an integer $s \geq 1$, the Sobolev class $H^s(\mc{X})$ consists of all those functions $f \in \Leb^2(\mc{X})$ which have weak derivatives of all orders $j = 1,\ldots,s$, all of which also belong to $\Leb^2(\mc{X})$.
-The $j$th order semi-norm for $f \in H^s(\mc{X})$ is $|f|_{H^j(\mc{X})} := \|d^jf\|_{L^2(Q)}$, and the corresponding norm
-\begin{equation*}
-\|f\|_{H^s(\mc{X})}^2 := \|f\|_{\Leb^2(Q)}^2 + \sum_{j = 1}^{s} |f|_{H^j(Q)}^2,
+\int_{\mc{X}} h g = (-1)^{|\alpha|} \int_{\mc{X}} f D^{\alpha}g, \quad \textrm{for every $g \in C_c^{\infty}(\mc{X})$.}
+\end{equation*}
+If such a function $h$ exists, it is the $\alpha$th weak partial derivative of $f$, and denoted by $D^{\alpha}f := h$.
+
+\begin{definition}[Sobolev space on an open set]
+	\label{def:sobolev_space}
+	For an integer $s \geq 1$, a function $f \in L^2(\mc{X})$ belongs to the Sobolev space $H^s(\mc{X})$ if for all $|\alpha| \leq s$, the weak derivatives $D^{\alpha}f$ exists and satisfy $D^{\alpha}f \in L^2(\mc{X})$.
+The $j$th order semi-norm for $f \in H^s(\mc{X})$ is $|f|_{H^j(\mc{X})} := \sum_{|j| = s}\|D^{\alpha}f\|_{\Leb^2(\mc{X})}$, and the corresponding norm
+	\begin{equation*}
+	\|f\|_{H^s(\mc{X})}^2 := \|f\|_{\Leb^2(\mc{X})}^2 + \sum_{j = 1}^{s} |f|_{H^j(\mc{X})}^2,
 	\end{equation*}
@@
 	\end{equation*} 
+\end{definition}
 We note that $H^s(\mc{X})$ is the completion of $C^{\infty}(\mc{X})$ with respect to the $\|\cdot\|_{H^s(\mc{X})}$ norm, so that $C^{\infty}(\mc{X})$ is dense in $H^s(\mc{X})$.
@@
 \paragraph{Manifold case.}
-There are several ways to define Sobolev spaces on a manifold \textcolor{red}{(references)}, but we will stick with a simple definition that suffices when $\mc{X}$ is a smooth Riemmanian manifold without boundary, as in Model~\ref{def:model_manifold}.
-Let $\{\phi_{\beta}, U_{\beta}\}$ be a $C^{\infty}$ atlas of $\mc{X}$, and let $\{h_\beta\}$ be a partition of unity subordinate to $\{(\phi_{\beta}), U_{\beta}\}$.
-For a function $f \in C^{\infty}(\mc{X})$, define the Sobolev norm to be,
-\begin{equation*}
-\|f\|_{H^s(\mc{X})}^2 = \sum_{\beta} \|(fh_{\beta}) \circ \phi_\beta^{-1}\|_{H^{k}(\Reals^m)}^2.
-\end{equation*}
-Then the order-$s$ Sobolev space $H^s(\mc{X})$ is defined to be the completion of $C^{\infty}(\mc{X})$ with respect to the norm $\|\cdot\|_{H^s(\mc{X})}$.
+
+There are several equivalent ways to define Sobolev spaces on a compact, smooth, $m$-dimensional Riemannian manifold embedded in $\Reals^d$.
+We will stick with a definition that parallels our setup in the flat Euclidean setting as much as possible.
+To do so, we first recall the notion of partial derivatives on a manifold, which are defined with respect to a local coordinate system.
+Letting $r_1,\ldots,r_m$ be the standard basis of $\Reals^m$, for a given chart $(\phi,U)$ (meaning an open set $U \subseteq \mc{X}$, and a smooth mapping $\phi: U \to \Reals^m$) we write $\phi =: (x_1,\ldots,x_m)$ in local coordinates, meaning $x_i = r_i \circ \phi$.
+Then the partial derivative $\partial f/\partial x_i$ of a function $f$ with respect to $x_i$ at $x \in U$ is
+\begin{equation*}
+\frac{\partial f}{\partial x_i}(x) := \frac{\partial(f \circ \phi^{-1})}{\partial r_i}\bigl(\phi(x)\bigr).
+\end{equation*}
+The right hand side should be interpreted in the weak sense of derivative.
+As before, we use the multi-index notation $D^{\alpha}f := \partial^{|\alpha|}f/\partial^{\alpha_1}x_1\ldots\partial^{\alpha_m}x_m$.
+
+\begin{definition}[Sobolev space on a manifold]
+	\label{def:sobolev_space_manifold}
+	A function $f \in \Leb^2(\mc{X})$ belongs to the Sobolev space $H^{s}(\mc{X})$ if for all $\abs{\alpha} \leq s$, the weak derivatives $D^{\alpha}f$ exist and satisfy  $D^{\alpha}f \in \Leb^2(\mc{X})$.
+The $j$th order semi-norm $|f|_{H^j(\mc{X})}$, the norm $\|f\|_{H^s(\mc{X})}$, and the ball $H^s(\mc{X};M)$ are all defined as in Defintion~\ref{def:sobolev_space}.
+\end{definition}
+The partial derivatives $D^{\alpha}f$ will depend on the choice of local coordinates, and so will the resulting Sobolev norm~$\|f\|_{H^s(\mc{X})}$.
+However, regardless of the choice of local coordinates the resulting norms will be equivalent\footnote{Recall that norms $\|\cdot\|_1$ and $\|\cdot\|_2$ on a space $\mc{F}$ are said to be equivalent if there exist constants $c$ and $C$ such that
+\begin{equation*}
+c \|f\|_1 \leq \|f\|_2 \leq C \|f\|_2 \quad \textrm{for all} f \in \mc{F}.
+\end{equation*}} 
+and so the ultimate Sobolev space $H^s(\mc{X})$ is independent of the choice of local coordinates.
+
+% AG: This was my first attempt at defining Sobolev spaces on manifolds, which I keep as a comment for now, and plan to delete.
+%
+% We stick to an approach that builds on the definition of Sobolev spaces in the flat Euclidean case.
+Specifically, let $\{\phi_{\beta}, U_{\beta}\}$ be a $C^{\infty}$ atlas of $\mc{X}$, and let $\{h_\beta\}$ be a partition of unity subordinate to $\{(\phi_{\beta}), U_{\beta}\}$.\footnote{Meaning $\mathrm{supp}(h_{\beta}) \subseteq U_{\beta}$ for each $\beta$.} For a function $f \in C^{\infty}(\mc{X})$, define the order-$s$ Sobolev norm to be
+%\begin{equation*}
+%\|f\|_{H^s(\mc{X})}^2 := \sum_{\beta} \|(fh_{\beta}) \circ %\phi_\beta^{-1}\|_{H^s(\phi(U_\beta))}^2.
+%\end{equation*}
+%Then the order-$s$ Sobolev space $H^s(\mc{X})$ is defined to be the completion of $C^{\infty}(\mc{X})$ with respect to the norm $\|\cdot\|_{H^s(\mc{X})}$.
+When the manifold $\mc{X}$ is smooth, as we assume in Model~\ref{def:model_manifold}, then this construction is independent of the choice of atlas $\{\phi_{\beta}, U_{\beta}\}$.
 
 \paragraph{Boundary conditions.}
-In the flat Euclidean case, when $s > 1$ we will assume that $f_0$ satisfies some boundary conditions.
-In particular, we will assume that $f_0$ belongs to the set of \emph{zero trace} Sobolev functions $H_0^{s}(P)$, defined as the closure of $C_0^s(\mc{X})$ with respect to $\|\cdot\|_{H^s(\mc{X})}$ norm.
-Of course, under Model~\ref{def:model_manifold} the domain $\mc{X}$ is without boundary, and so boundary conditions are irrelevant.
-
-Let us explain the reason we require boundary conditions in the flat Euclidean case when $s > 1$, by considering an alternative way to characterize Sobolev spaces using a relation to sequence spaces.\footnote{This is equivalent to characterizing Sobolev spaces using the Fourier transform, but we find it more transparent to instead work directly with sequence spaces.} Let $(\lambda_1(\Delta_P),\psi_1),(\lambda_2(\Delta_P),\psi_2),\ldots$ be the solutions to the Laplace-Beltrami eigenvector equation (with Neumann boundary conditions),
-\begin{equation*}
-\Delta_P\psi_k = \lambda_{k}(\Delta_P) \psi_k, \quad \frac{\partial}{\partial {\bf n}}\psi_k = 0~~\textrm{on $\partial \mc{X}$},
-\end{equation*}
-where $\partial/(\partial {\bf n})$ is the partial derivative operator in the direction of the normal vector $\mathbf{n}$ to the boundary of $\mc{X}$.
-We define the ellipsoid $\Theta \subset \ell^2(\mathbb{N})$  using the eigenvalues $\lambda_1(\Delta_P),\lambda_2(\Delta_P),\ldots$ to be
-\begin{equation*}
-\Theta := \Bigl\{\theta \in \ell^2(\mathbb{N}): \sum_{k = 1}^{\infty} [\lambda_k(\Delta_P)]^{2s} \theta_k^2 < \infty\Bigr\}
-\end{equation*}
-Then, the function class $\mc{H}^s(P) = \{f = \sum_{k = 1}^{\infty} \theta_k \psi_k: \theta \in \Theta\}$ consists precisely of those Sobolev functions $f \in H^s(\mc{X})$ which satisfy the following \emph{Neumann} boundary conditions,
-\begin{equation}
-\label{eqn:condition}
-\frac{\partial^j}{\partial {\bf n}^j}f = 0 \quad \textrm{for $j = 1,3,5,\ldots,\floor{s/2}$.}
-\end{equation}
-As we have already reviewed, the eigenvectors $v_k$ of the graph Laplacian $L_{n,\varepsilon}$ converge to $\psi_k$ as $n \to \infty, \varepsilon \to 0$.
-Naturally, then, these eigenvectors accurately approximate only those functions $f_0 \in \mc{H}^s(P) \subset H^s(\mc{X})$.
-Although the zero-trace boundary condition is more restrictive than~\eqref{eqn:condition}---since it also requires that derivatives of even-order be equal to $0$---the point is that some kind of boundary condition will be necessary in order to obtain optimal rates for spectral series methods, including Laplacian eigenmaps.
+In the flat Euclidean model (Model~\ref{def:model_flat_euclidean}), in order to show that Laplacian eigenmaps is optimal over $H^{s}(\mc{X})$ for $s > 1$ we will need to assume that the regression function $f_0$ satisfies some boundary conditions.
+In particular, we will assume that $f_0$ is zero-trace.
+\begin{definition}[Zero-trace Sobolev space]
+	\label{def:zero_trace_sobolev_space}
+	The \emph{order-s zero-trace Sobolev space} $H_0^s(\mc{X})$ is the closure of $C_c^\infty(\mc{X})$ with respect to $\|\cdot\|_{H^s(\mc{X})}$ norm.
+That is, $f \in H_0^s(\mc{X})$ if $f \in H^s(\mc{X})$ and additionally there exists a sequence $f_1,f_2,\ldots$ of functions in $C_c^{\infty}(\mc{X})$ such that
+	\begin{equation*}
+	\lim_{k \to \infty}\|f_k - f\|_{H^s(\mc{X})} = 0.
+	\end{equation*}
+	The normed ball $H_0^{s}(\mc{X};M) := H_0^{s}(\mc{X}) \cap H^{s}(\mc{X};M)$.
+\end{definition}
+The zero-trace condition can be made more concrete when $f \in C^\infty(\mc{X})$, since we can then speak of the pointwise behavior of $f$ and its derivatives.
+Letting $\partial/(\partial {\bf n})$ be the partial derivative operator in the direction of the vector $\mathbf{n}$ normal to the boundary of $\mc{X}$, then the zero-trace condition implies that $\partial^{k}f/\partial{\bf n}^k(x) = 0$ for each $k = 0,\ldots,s - 1$, and for all $x \in \partial\mc{X}$.
+
+We now explain why Laplacian eigenmaps should be optimal only when $f_0$ satisfies certain boundary conditions.
+Let $(\lambda_1(\Delta_P),\psi_1),(\lambda_2(\Delta_P),\psi_2),\ldots$ be the solutions to the weighted Laplacian eigenvector equation with Neumann boundary conditions,
+\begin{equation}
+\label{eqn:laplacian_eigenfunction_neumann}
+\Delta_P\psi_k = \lambda_{k}(\Delta_P) \psi_k, \quad \frac{\partial}{\partial {\bf n}}\psi_k = 0~~\textrm{on $\partial \mc{X}$}.
+\end{equation}
+As we have already alluded to, it is known~\citep{garciatrillos18} that each graph Laplacian eigenpair ($\lambda_k$,$v_k$) converges to a corresponding solution ($\lambda_k(\Delta_P)$, $\psi_k$) of~\eqref{eqn:laplacian_eigenfunction_neumann}.
+Thus it is relevant to consider which Sobolev functions $f \in H^s(M)$ we can reconstruct using the eigenfunctions $\psi_1,\psi_2,\ldots$.
+To that end, we introduce the spectrally defined Sobolev space
+\begin{equation}
+\label{eqn:spectral_sobolev_space}
+\mc{H}^s(\mc{X}) := \biggl\{f \in L^2(\mc{X}): \sum_{k = 1}^{\infty} \bigl[\dotp{f}{\psi_k}_P\bigr]^2 \cdot\bigl[\lambda_k(\Delta_P)\bigr]^s < \infty \biggr\}.
+\end{equation}
+Under the conditions $p \in C^{\infty}(\mc{X})$ and $\partial\mc{X} \in C^{1,1}$, \cite{dunlop2020} show the strict inclusion $\mc{H}^{2s}(\mc{X}) \subset H^{2s}(\mc{X})$.
+More precisely, they show that for any $s > 0$,
+\begin{equation}
+\label{eqn:spectral_sobolev_space_equivalence}
+\mc{H}^{2s}(\mc{X}) = \biggl\{f \in H^{2s}(\mc{X}): \frac{\partial \Delta_P^rf}{\partial {\bf n}} = 0~\textrm{on}~\partial\mc{X},~~\textrm{for all $0 \leq r \leq s - 1$} \biggr\},
+\end{equation}
+and for any $s \geq 0$, $\mc{H}^{2s + 1}(\mc{X}) = \mc{H}^{2s}(\mc{X}) \cap H^{2s + 1}(\mc{X})$.
+If $P$ is uniform on $\mc{X}$, then~\eqref{eqn:spectral_sobolev_space} means that a Sobolev function $f \in H^{s}(\mc{X})$ additionally belongs to $\mc{H}^{s}(\mc{X})$ only if all its odd lower order derivatives vanish at the boundary $\partial\mc{X}$.
+
+The bottom line is that the eigenvectors $v_k$ of the graph Laplacian accurately approximate only those functions $f \in H^{s}(\mc{X})$ which satisfy some additional boundary conditions.
+Although the zero-trace boundary condition is more restrictive than~\eqref{eqn:spectral_sobolev_space_equivalence}---since it also requires that derivatives of even-order be equal to $0$---the point is that some kind of boundary condition will be necessary in order to obtain optimal rates.
 We will stick to the zero-trace condition, since it greatly eases some of the steps in our proofs.
 
-\subsection{Minimax Rates}
+On the other hand, in the manifold model (Model~\ref{def:model_manifold}) the domain $\mc{X}$ is without boundary---precisely, every point $x \in \mc{X}$ has a neighborhood that is homeomorphic to an open set in $\Reals^m$, for instance the open ball $B(x,\delta)$ for any $\delta$ smaller than the injectivity radius $i_0$---and so boundary conditions are irrelevant.
+
+\subsection{Minimax Rates and Spectral Series Methods}
 \label{subsec:minimax_rates_sobolev}
 We now review the minimax estimation and goodness-of-fit testing rates over Sobolev balls.
-We pay special attention to the spectral series methods which achieve these rates, because, as we have already seen, these methods are very related to Laplacian eigenmaps.
+We will pay special attention to certain classical spectral projection methods which achieve these rates.
+This is because, as we have already discussed, classical spectral projection methods are very related to Laplacian eigenmaps.
 
 \paragraph{Estimation rates over Sobolev balls.}
-In the estimation problem, we ask for an estimator---formally, a Borel measurable function $\wh{f}$, that maps from $\mc{X}$ to $\Reals$---which is close to the regression function $f_0$ with respect to the squared norm $\|\wh{f} - f_0\|_{P}^2$.
-Under Model~\ref{def:model_flat_euclidean}, the minimax estimation rate over the Sobolev ball is\footnote{Perhaps surprisingly, we were unable to find a reference that explicitly states the rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate} in our general setup, for all values of $s$ and $d$---but the rate certainly does hold.
-Momentarily we will introduce a spectral series estimator $\wt{f}$ which achieves this rate, implying the upper bound.
-The lower bound in~\eqref{eqn:sobolev_space_minimax_estimation_rate} can be obtained from the analogous lower bound over $C_0^s(\mc{X})$\textcolor{red}{(reference)}, and the fact that there exists a continuous embedding of $C_0^s(\mc{X})$ into $\mc{H}^s(P)$.}
+In the estimation problem, we ask for an estimator---formally, a Borel measurable function $\wh{f}$ that maps from $\mc{X}$ to $\Reals$---which is close to the regression function $f_0$ with respect to the squared norm $\|\wh{f} - f_0\|_{P}^2$.
+Under Model~\ref{def:model_flat_euclidean}, the minimax estimation rate over the order-$s$ Sobolev ball is
 \begin{equation}
 \label{eqn:sobolev_space_minimax_estimation_rate}
-\inf_{\wh{f}} \sup_{f_0} \|\wh{f} - f_0\|_P^2 \asymp M^2(M^2n)^{-2s/(2s + d)}
-\end{equation}
-here the infimum is taken over all estimators $\wh{f}$, and the supremum over all $f_0 \in H^1(\mc{X};M)$ (first-order case) or $f_0 \in H_0^s(\mc{X};M)$ (higher-order case).
-
-The upper bound in~\eqref{eqn:sobolev_space_minimax_estimation_rate} can be certified by a particular spectral series estimator.
-Under the assumptions of Model~\ref{def:model_flat_euclidean}, the eigenvalues $\lambda_k(\Delta_P)$ satisfy the Weyl's Law asymptotic scaling $\lambda_{k}(\Delta_P) \asymp k^{2/d}$, an important fact that we will make use of and discuss in more detail later on.
-A key consequence of this scaling is that the Sobolev ball $H_0^s(\mc{X}) \subset \mc{H}^s(P)$ continuously embeds into the Sobolev ellipsoid $\mc{\wt{H}}^s(P) = \{f = \sum_{k = 1}^{\infty} \theta_k \psi_k: \sum_{k = 1}^{\infty} (k - 1)^{2s/d} \theta_k^2\}$.
-By reducing the regression problem to an estimation problem in a Gaussian sequence model \textcolor{red}{(see e.g.
-Tsybakov, Johnstone, Gine)}, it can be shown the estimator
+\inf_{\wh{f}} \sup_{f_0} \Ebb\|\wh{f} - f_0\|_P^2 \asymp M^2(M^2n)^{-2s/(2s + d)};
+\end{equation}
+here infimum taken over all estimators $\wh{f}$, and the supremum over all $f_0 \in H^1(\mc{X};M)$ (first-order case) or $f_0 \in H_0^s(\mc{X};M)$ (higher-order case), and we assume $n^{-1/2} \lesssim M$.
+
+The lower bound in~\eqref{eqn:sobolev_space_minimax_estimation_rate} is due to~\citep{stone1980} (at least for the case of $M$ constant, as is most typically considered).
+The upper bound can be certified by a particular spectral projection estimator $\wt{f}$,
 \begin{equation}
@@
 \end{equation}
-achieves the minimax rate over normed balls in $\mc{\wt{H}}^s(P)$, and the same upper bound on the minimax rate therefore applies to normed balls $H_0^s(\mc{X};M)$ as well.
+where $\psi_k$ are the eigenfunctions of $\Delta_P$ defined in~\eqref{eqn:laplacian_eigenfunction_neumann}.
+The optimality of spectral projection estimators over Sobolev type spaces is generally well-understood.
+For instance, see \cite{tsybakov08,johnstone2011,gine16}, who work in the Gaussian sequence model and show that analogous estimators are optimal over Sobolev ellipsoids.
+However, we have not found an analysis of the specific estimator~\eqref{eqn:spectral_series_estimator} under Model~\ref{def:model_flat_euclidean}, and so for completeness we state the result in the following proposition.
+\begin{proposition}
+	\label{prop:spectral_series_estimation}
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $\partial \mc{X} \in C^{1,1}$, that $p \in C^{\infty}(\mc{X})$, and that $f_0 \in H^1(\mc{X};M)$ (first-order case) or $f_0 \in H_0^s(\mc{X};M)$ for some $s > 1$ (higher-order case).
+Then there exists a constant $C$ which does not depend on $f_0,M$ or $n$ such that the following statement holds: if the spectral projection estimator $\wt{f}$ is computed with parameter $K = \floor{M^2n}^{d/(2s + d)}$, then
+	\begin{equation*}
+	\Ebb\bigl[\|\wt{f} - f_0\|_P^2\bigr] \leq C \min\bigl\{M^2(M^2n)^{-2s/(2s + d)}, M^2\bigr\}
+	\end{equation*}
+\end{proposition}
+The proof of Proposition~\ref{prop:spectral_series_estimation}, along with proofs of all of our results, can be found in the appendix.
+It is worth mentioning some aspects of the analysis here, because it sets the stage for the strategy we will use to analyze Laplacian eigenmaps.
+
+In particular, three essential facts are needed to establish Proposition~\ref{prop:spectral_series_estimation}. 
+\begin{enumerate}
+	\item The continuous embedding of $H_0^s(\mc{X})$ into $\mc{H}^s(\mc{X})$---recall the latter is defined in~\eqref{eqn:spectral_sobolev_space}---which is a consequence of the zero-trace condition, the conditions on $\partial \mc{X}$ and $p$, and~\eqref{eqn:spectral_sobolev_space_equivalence}.
+	\item Weyl's Law, which gives the asymptotic scaling of eigenvalues $\lambda_{k}(\Delta_P) \asymp k^{2/d}$, and allows us to properly control the bias induced by spectral projection.
+	\item A local version of Weyl's law, which gives an estimate on $\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2$, and allows us to appropriately control the difference $\dotp{f_0}{\psi_k}_n - \dotp{f_0}{\psi_k}_P$  between the empirical and population Fourier coefficients.
+\end{enumerate}
+With these facts in hand the proof of Proposition~\ref{prop:spectral_series_estimation} follows from calculations standard to analysis of the Gaussian sequence model.
+Our analysis of Laplacian eigenmaps will depend on analogues to the first and second of these facts, with the space $\mc{H}^s(\mc{X})$ and eigenvalues $\lambda_k(\Delta_P)$ replaced by alternatives suitably defined with respect to the neighborhood graph Laplacian $L_{n,\varepsilon}$.
 
@@
 \begin{equation}
-\mathbf{H}_0: f_0 = f_0^{\star}, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{f_0^{\star}\}.
-\end{equation} 
-Typically, the null hypothesis $f_0 = f_0^{\star} \in \mc{F}$ reflects the absence of interesting structure, and $\mc{F} \setminus  \{f_0^{\star}\}$ is a set of smooth departures from this null.
-To fix ideas, as in \citet{ingster2009} we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\star} = 0$ and $\mc{F} = \mc{H}^s(\mc{X};M)$ is a Sobolev ball.
-This is without loss of generality since our test statistic and its analysis are easily modified to handle the case when $f_0^{\star}$ is not $0$, by simply subtracting $f_0^{\star}(X_i)$ from each observation $Y_i$.
+\mathbf{H}_0: f_0 = f_0^{\ast}, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{f_0^{\ast}\}.
+\end{equation} 
+Typically, the null hypothesis $f_0 = f_0^{\ast} \in \mc{F}$ reflects the absence of interesting structure, and $\mc{F} \setminus  \{f_0^{\ast}\}$ is a set of smooth departures from this null.
+To fix ideas, as in \citet{ingster2009} we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\ast} = 0$ and $\mc{F} = \mc{H}^s(\mc{X};M)$ is a Sobolev ball.
+This is without loss of generality since our test statistic and its analysis are easily modified to handle the case when $f_0^{\ast}$ is not $0$, by simply subtracting $f_0^{\ast}(X_i)$ from each observation $Y_i$.
 
@@
 \end{equation*}
-and for a given constant $b \in (0,1)$, the minimax critical radius $\epsilon(\mc{F})$ is the smallest value of $\epsilon$ such that some level-$a$ test has worst-case risk of at most $b$. Formally,
+and for a given constant $b \in (0,1)$, the minimax critical radius $\epsilon_n(\mc{F})$ is the smallest value of $\epsilon$ such that some level-$a$ test has worst-case risk of at most $b$. Formally,
 \begin{equation*}
@@
 \end{equation*} 
-where in the above the infimum is over all level-$a$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Clearly, the minimax critical radius $\epsilon(\mc{F})$ depends on $a$ and $b$.
-However, we adopt the typical convention of treating $\alpha,b \in (0,1)$ and  as small but fixed positive constants; hence they will not affect the testing error rates, and we suppress them notationally.} See \citet{ingster82,ingster87,ingster2012} for a more extended treatment of the minimax paradigm in nonparametric testing.
-
-Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing critical radius is much smaller than the minimax estimation rate.
-\citet{ingster2009} show that in the special case where $\mc{X} = [0,1]^d$ and $P = \nu$,
+where in the above the infimum is over all level-$a$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Clearly, the minimax critical radius $\epsilon_n(\mc{F})$ depends on $a$ and $b$.
+However, we adopt the typical convention of treating $\alpha,b \in (0,1)$ and  as small but fixed positive constants; hence they will not affect the testing error rates, and we suppress them notationally.} We will refer to the rate at which the squared critical radius $\epsilon_n(\mc{F})^2$ tends to $0$ as the minimax testing rate.
+
+Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing rate is much smaller than the minimax estimation rate.
+\citet{ingster2009} give the minimax testing rate in the special case where $\mc{X} = [0,1]^d$ and $P$ is uniform, and we restate their result in the terms and notation of our paper.
+\begin{theorem}[Theorem 1 of \citet{ingster2009}]
+	\label{thm:spectral_series_testing}
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $\mc{X} = [0,1]^d$, $P$ is the uniform distribution on $[0,1]^d$, and $f_0 \in H_0^s(\mc{X})$. Then 
 	\begin{equation}
 	\label{eqn:sobolev_space_testing_critical_radius}
-\epsilon_n^2\bigl(\mc{H}^s(P;M)\bigr) \asymp M^2(M^2n)^{-4s/(4s + d)}~~\textrm{for $1 \leq d < 4s$,}
-\end{equation}
-and the same result holds under the more general setup of Model~\ref{def:model_flat_euclidean}.
-The test \citet{ingster2009} use to certify the upper bound in~\eqref{eqn:sobolev_space_testing_critical_radius} is a spectral series method, which uses the squared $L^2(P)$ norm of $\wt{T} = \|\wt{f}\|_{P}^2$ as a test statistic.
-
-When $4s \geq d$ the functions in $H^s(\Xset)$ are very irregular---formally speaking $H^s(\Xset)$ does not continuously embed into $\Leb^4(\Xset)$ when $4s \geq d$---and as far as we know, the minimax testing rates in this regime have not been worked out.
-However, if one explicitly assumes that $f_0 \in \Leb^4(\mc{X})$ then the critical radius is given by $\epsilon_n^2(\Leb^4(\mc{X};M)) \asymp Mn^{-1/2}$.
-As a sanity check note that this is strictly worse than the rate in~\eqref{eqn:sobolev_space_testing_critical_radius}.
-As we discuss after our first main theorem regarding testing (Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}), this rate is achievable by a test based on $\wh{T}$.
+	\epsilon_n^2\bigl(H_0^s(P;1)\bigr) \asymp n^{-4s/(4s + d)}~~\textrm{for $1 \leq d < 4s$.}
+	\end{equation}
+\end{theorem}
+The analysis used by \citet{ingster2009} to show the upper bound in~\eqref{eqn:sobolev_space_testing_critical_radius} relies on a similar trio of facts as used in the proof of Proposition~\ref{prop:spectral_series_estimation}.
+It is otherwise reminiscent of calculations made in the Gaussian sequence model, which can be found in~\cite{ingster2012}.
+This analysis can be straightforwardly adapted to handle design distributions $P$ that satisfy the conditions of Model~\ref{def:model_flat_euclidean}, or to handle the case where $M$ is not $1$.
+Finally, the test \citet{ingster2009} use to certify the upper bound in~\eqref{eqn:sobolev_space_testing_critical_radius} is implicitly a spectral projection method: the test statistic is the $L^2(P_n)$ norm of a particular spectral projection estimator.\footnote{\cite{ingster2009} project the responses onto a the span of trigonometric basis functions.
+Since $P$ is uniform on the unit cube, such functions are eigenfunctions of $\Delta_P$ when the right boundary conditions are imposed.}
+
+A major difference between testing and estimation over Sobolev spaces is the requirement that $4s > d$.
+When $4s \leq d$, the functions in $H^s(\Xset)$ are very irregular.
+Crucially, $H^s(\Xset)$ no longer continuously embeds into $\Leb^4(\Xset)$ when $4s \leq d$, and test statistics using the $L^2(P_n)$ norm are no longer guaranteed to have finite variance.\footnote{Note that this will not affect the analysis for estimation, because for estimation we only need to control the first two moments of $f_0$.} In fact, we have not seen any analysis which describes the minimax rate for nonparametric regression testing over Sobolev spaces in the $4s \leq d$ regime.
+However, if one explicitly assumes that $f_0 \in \Leb^4(\mc{X})$, then the critical radius is characterized by the dimension-free rate $\epsilon_n^2(\Leb^4(\mc{X};M)) \asymp Mn^{-1/2}$.\footnote{As a sanity check note that this is strictly worse than the rate in~\eqref{eqn:sobolev_space_testing_critical_radius}.
+In other words, whenever $4s > d$, the embedding $H^s(\mc{X}) \subseteq L^4(\mc{X})$ never yields a tight upper bound.} As we discuss after our first main theorem regarding testing with Laplacian eigenmaps (Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}), this rate is achievable by a test based on $\wh{T}$.
 
@@
 Under Model~\ref{def:model_manifold}, both~\eqref{eqn:sobolev_space_minimax_estimation_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced everywhere by the intrinsic dimension $m$.
-The estimator $\wt{f}$ and a test using the statistic $\wt{T}$---with $\psi_1,\psi_2,\ldots$ now the eigenfunctions of the manifold weighted Laplace-Beltrami operator $\Delta_P$---continue to achieve the optimal rates.
-\textcolor{red}{(TODO): Find references.}
+The estimator $\wt{f}$ and a test using the statistic $\wt{T}$---with $\psi_1,\psi_2,\ldots$ now the eigenfunctions of the manifold weighted Laplace-Beltrami operator $\Delta_P$---achieve the optimal estimation and testing rates.
+This is because each of the three facts mentioned after Proposition~\ref{prop:spectral_series_estimation} have analogues when the domain is a smooth manifold (See~\citet{hendriks1990}, who analyzes a spectral projection density estimator, for details.)
 
 \paragraph{In-sample mean squared error.}
-
-The error in~\eqref{eqn:sobolev_space_minimax_estimation_rate} is measured with respect to $\|\cdot\|_P^2$ norm.
-Since $\wh{f}$ is defined only at the design points $X_1,\ldots,X_n$, we will measure estimation rates in $\|\cdot\|_n^2$ error.
-The lower bounds used to show~\eqref{eqn:sobolev_space_minimax_estimation_rate} will still hold when in-sample loss is used rather than out-of-sample loss \textcolor{red}{(TODO): (I don't actually know if this is true.
-Need to check.)}, and we verify matching upper bounds in the sections to come.
-Thus the minimax rates are unchanged---we call estimators which achieve these optimal rates \emph{in-sample minimax rate-optimal estimators}.
+As mentioned in our introduction, roughly speaking one of our main conclusions is that the Laplacian eigenmaps estimator $\wh{f}$ is minimax rate-optimal.
+It is worth being clear about what we do and do not mean by this statement.
+We do not mean that the estimator $\wh{f}$ will match the upper bound given in~\eqref{eqn:sobolev_space_minimax_estimation_rate}, since such a statement does not make sense when the estimator is defined only at the random design points $X_1,\ldots,X_n$.
+Instead we will measure loss using the squared $L^2(P_n)$ error.
+In Section~\ref{sec:out_of_sample} we show that an extension of $\wh{f}$ defined over all $\mc{X}$ has $L^2(P)$ error comparable to the $L^2(P_n)$ error of $\wh{f}$.
+We also believe that in the random design setting we work in, simple arguments will imply that $L^2(P_n)$ risk has the same minimax rate of convergence as $L^2(P)$ risk.
+We sketch such an argument in Section~\ref{sec:out_of_sample}, but do not further pursue the details.
+
+Additionally, we will not actually measure accuracy using the expectation of the loss.
+Rather, we will give a high-probability bound on $\|\cdot\|_n^2$.
+For instance, when $f_0 \in H^1(\mc{X};1)$, we will show that with probability $1 - \delta$ the loss $\|\wh{f} - f_0\|_n^2 \leq C_{\delta} n^{-2/(2 + d)}$, for a constant $C_{\delta}$ that depends on $\delta$ but not on $f_0$ or $n$.
+Thus we give an upper bound on the $(1 - \delta)$th quantile of $\|\wh{f} - f_0\|_n^2$, rather than an upper bound on its expectation.
+We explain the reason for this in Section~\ref{sec:minimax_optimal_laplacian_eigenmaps}.
+We also show that if $f_0$ is bounded in a larger norm---for instance, if it is H\"{o}lder rather than Sobolev smooth---then we can obtain bounds on the expected $L^2(P_n)$ loss.
 
@@
 Thus one cannot speak of the pointwise evaluation $f_0(X_i)$, as we do by defining our target of estimation to be $f_0(X_i)$, $i=1,\ldots,n$, until one selects \emph{representatives}.
-When $s > d/2$, every element $f$ of $H^s(\mc{X})$ admits a continuous version $f^{\ast}$, and it is standard to pick this as the favored representative.
+When $s > d/2$, every element $f$ of $H^s(\mc{X})$ admits a continuous version $f^{\ast}$, and as is standard we set this to be our favored representative.
 When $s \leq d/2$, some elements in $H^s(\mc{X})$ do not have any continuous version; however they admit a \emph{quasi-continuous} version \citep{evans15} known as the \emph{precise representative}, and we use this representative.
@@
 
-% TODO: AG, get rid of this!
-
-%\subsection{Our contributions}
-%The following items summarize our major results.
-In all of them we assume~\ref{asmp:domain} and~\ref{asmp:density} hold, but not that any of $P$, $\mc{X}$, $\Delta_P$ or $\Delta$ are known.
-Put $r := s - 1$.
-%\begin{itemize}
-%	\item \textbf{Minimax optimal estimation.} For any $f_0 \in H^1(\Xset;M)$, with high probability $\norm{\wh{f} - f_0}_{n}^2 \lesssim M^{2d/(2 + d)}n^{-2/(2 + d)}$.
-%	\item \textbf{Minimax optimal testing.}
-%	A test constructed using $\wh{T}$ has non-trivial power whenever $f_0 \in H^1(\Xset;M)$ satisfies $\norm{f_0}_{P}^2 \gtrsim M^{2d/(2 + d)}n^{-4/(4 + d)}$ and $d < 4$.
-%	\item \textbf{Higher-order smoothness conditions.} When the regression function $f_0 \in H_0^s(\Xset;M)$ and additionally the density $p \in C^{r}(\mc{X};M)$, with high probability the error $\norm{\wh{f} - f_0}_{n}^2 \lesssim M^{2d/(2s + d)}n^{-2s/(2s + d)}$.
-Additionally, a test constructed using $\wh{T}$ has non-trivial power whenever $f_0 \in H_0^s(\Xset;M)$ satisfies $\norm{f_0}_{P}^2 \gtrsim M^{2d/(4s + d)}n^{-4s/(4s + d)}$ and $d < 4$.
-%	\item \textbf{Manifold adaptivity.}
-%	If $\mc{X} \subset \Rd$ is a manifold of dimension $m < d$, and if $H^s(\mc{X})$ for $s \leq 4$, then each of the aforementioned rates hold with $d$ replaced by $m$.
-%\end{itemize}
-%In the statements above, estimation error is measured in-sample, using the empirical squared-norm $\|\cdot\|_n^2$.\footnote{To state the obvious: this is very different than measuring the \emph{training error} $\|Y - \wh{f}\|_n^2$.} It is necessary to measure loss in-sample because the Laplacian eigenmaps estimator $\wh{f}$ is defined only at the observed design points $X_1,\ldots,X_n$.
-That being said, often one would like an upper bound on the prediction risk, or equivalently the out-of-sample loss $\|\cdot\|_P^2$.
-We therefore also introduce a simple, intuitive way of extending $\wh{f}$ to a function defined over all of $\mc{X}$, and show that this extension is minimax rate-optimal with respect to error measured in $L^2(P)$ norm.
+Finally, we note that for testing none of these comments are relevant.
+We will show that our test has small worst case risk whenever $\epsilon \gtrsim \epsilon_n(H_0^s(\mc{X};M))$, thus establishing that it is a minimax optimal test in the usual sense.
 
@@
 
-As previously explained, Laplacian eigenmaps is a discrete and noisy approximation to a spectral series method, of particular use when the design distribution $P$ is unknown.
-Our goal is to show that the Laplacian eigenmaps methods are rate optimal, notwithstanding the extra error incurred by this approximation.
+As previously explained, Laplacian eigenmaps is a discrete and noisy approximation to a spectral projection method using the eigenfunctions of $\Delta_P$.
+This is particularly useful when $P$ is unknown, or when the eigenfunctions of $\Delta_P$ cannot be explicitly computed.
+Our goal is to show that Laplacian eigenmaps methods are rate-optimal, notwithstanding the potential extra error incurred by this approximation.
 In this section and the following one, we will see that this is indeed the case: the estimator $\wh{f}$, and a test using the statistic $\wh{T}$, achieve optimal estimation and goodness-of-fit testing rates over Sobolev classes.
 
-In this Section we will cover the flat Euclidean case, where we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to Model~\ref{def:model_flat_euclidean}.
+In this section we will cover the flat Euclidean case, where we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to Model~\ref{def:model_flat_euclidean}.
 We will divide our theorem statements based on whether we assume the regression function $f_0$ belongs to the first order Sobolev class ($s = 1$) or a higher-order Sobolev class ($s > 1$), since the details of the two settings are somewhat different.
@@
 
-\paragraph{Estimation.} When the kernel $\eta$, graph radius $\varepsilon$, and number of eigenvectors $K$ are chosen appropriately, we show in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that the estimator $\wh{f}$ achieves the minimax rate over $H^1(\mc{X};M)$, when error is measured in $\|\cdot\|_{n}$.
-\begin{enumerate}[label=(A\arabic*)]
+\paragraph{Estimation.} When the kernel $\eta$, graph radius $\varepsilon$, and number of eigenvectors $K$ are chosen appropriately, we show in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that the estimator $\wh{f}$ achieves the minimax rate over $H^1(\mc{X};M)$, when error is measured in squared $L^2(P_n)$ norm.
+\begin{enumerate}[label=(K\arabic*)]
 	\setcounter{enumi}{0}
@@
 	\end{equation*}
-	and we assume \smash{$\int_{\Rd} \|x\|^2 \eta(\|x\|) \,dx < \infty$}.
+	and we assume \smash{$\sigma_{\eta} := \frac{1}{d}\int_{\Rd} \|x\|^2 \eta(\|x\|) \,dx < \infty$}.
+\end{enumerate}
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{0}
 	\item 
 	\label{asmp:parameters_estimation_fo} 
-	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy the following inequalities:
+	For constants $c_0$ and $C_0$, the graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy the following inequalities:
 	\begin{equation}\\
 	\label{eqn:radius_fo} 
-	\biggl(\frac{\log n}{n}\biggr)^{1/d} \ll \varepsilon \ll \min\{1,K^{-1/d}\},
+	C_0\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq \varepsilon \leq c_0\min\{1,K^{-1/d}\},
 	\end{equation}
@@
 	\label{eqn:eigenvector_estimation_fo} 
-	K \asymp \min\Bigl\{(M^2n)^{d/(2 + d)} \vee 1, n\Bigr\}
+	K = \min\Bigl\{\floor{(M^2n)^{d/(2 + d)}} \vee 1, n\Bigr\}.
 	\end{equation}
 \end{enumerate}
+We comment on these assumptions after stating our first main theorem, regarding the estimation error of Laplacian eigenmaps.
 \begin{theorem}
@@
 	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H^1(\mc{X},M)$.
-There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_fo}, then
+There are constants $c,C$ and $N$ (not depending on $f_0$, $M$ or $n$), such that the following statement holds for all $n \geq N$ and any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_fo}, then
 	\begin{equation}
@@
 	\end{equation}
-	with probability at least $1 - \delta - Cn\exp(-cn\varepsilon^d)$.
+	with probability at least $1 - \delta - Cn\exp(-cn\varepsilon^d) - \exp(-K)$.
 \end{theorem}
-From~\eqref{eqn:laplacian_eigenmaps_estimation_fo} it follows immediately that with high probability $\|\wh{f} - f_0\|_n^2 \lesssim M^2(M^2n)^{-2/(2 + d)}$ whenever $n^{-1/2} \lesssim M \lesssim n^{1/d}$, and is thus an in-sample minimax rate-optimal estimator.
+From~\eqref{eqn:laplacian_eigenmaps_estimation_fo} it follows immediately that with high probability $\|\wh{f} - f_0\|_n^2 \lesssim M^2(M^2n)^{-2/(2 + d)}$ whenever $n^{-1/2} \lesssim M \lesssim n^{1/d}$.
 
@@
 \begin{itemize}
-	\item When $M \ll n^{-1/2}$ or $M \gg n^{1/d}$, the rate in~\eqref{eqn:laplacian_eigenmaps_estimation_fo} looks different than~\eqref{eqn:sobolev_space_minimax_estimation_rate}.
-In particular, when $M \ll n^{-1/2}$, then computing Laplacian eigenmaps with $K = 1$ achieves the parametric rate $\|\wh{f} - f_0\|_n^2 \lesssim n^{-1}$, and the zero-estimator $\wh{f} = 0$ achieves the better rate $\|\wh{f} - f_0\|_n^2 \lesssim M^2$, but neither match the minimax rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}.
-On the other hand, when $M \gg n^{1/d}$, then computing Laplacian eigenmaps with $K = n$ achieves the rate $\|\wh{f} - f_0\|_n^2 \asymp 1$, which is better than the rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}.
+	\item When $M = o(n^{-1/2})$, then computing Laplacian eigenmaps with $K = 1$ achieves the parametric rate $\|\wh{f} - f_0\|_n^2 \lesssim n^{-1}$, and the zero-estimator $\wh{f} = 0$ achieves the better rate $\|\wh{f} - f_0\|_n^2 \lesssim M^2$.
+However, we do not know what the minimax rate is in this regime.
+On the other hand, when $M = \omega(n^{1/d})$, then computing Laplacian eigenmaps with $K = n$ achieves the rate $\|\wh{f} - f_0\|_n^2 \lesssim 1$, which is better than the rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}.
+This is because we are evaluating error in $L^2(P_n)$ rather than $L^2(P)$.
 However, in truth these are edge cases, which do not fall neatly into the framework of nonparametric regression.
-	\item The lower bound on $\varepsilon$ imposed by~\eqref{eqn:radius_fo} is on the order of the connectivity threshold, the smallest length scale at which the resulting graph will still, with high probability, be connected.
-On the other hand, as we will see in Section~\ref{subsec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K$ is on the same order as the continuum eigenvalue $\lambda_K(\Delta_P)$.
-Finally, we choose $K \asymp (M^2n)^{2d/(2 + d)}$ to optimally trade-off bias and variance.
-	\item We note that the ranges~\eqref{eqn:radius_fo} and~\eqref{eqn:eigenvector_estimation_fo} depend on unknown quantities such as the dimension $d$ and radius of the Sobolev ball $M$.
+	\item The assumptions placed on the kernel function $\eta$ are needed for technical reasons.
+They can likely be weakened, although we note that they are already fairly general.
+The lower bound on $\varepsilon$ imposed by~\eqref{eqn:radius_fo} is on the order of the connectivity threshold, the smallest length scale at which the resulting graph will still be connected with high probability.
+On the other hand, as we will see in Section~\ref{subsec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K$ is of at least the same order as the continuum eigenvalue $\lambda_K(\Delta_P)$.
+Finally, we choose $K = (M^2n)^{2d/(2 + d)}$ (when possible) to optimally trade-off bias and variance.
+	\item We note that the ranges~\eqref{eqn:radius_fo} and~\eqref{eqn:eigenvector_estimation_fo} depend on quantities, such as the dimension $d$ and radius of the Sobolev ball $M$, which are usually unknown.
 In practice, one typically tunes hyper-parameters by sample-splitting or cross-validation.
@@
 We return to this issue in Section~\ref{sec:out_of_sample}, when we propose an out-of-sample extension of $\wh{f}$.
-	\item The upper bound given in Equation~\eqref{eqn:laplacian_eigenmaps_estimation_fo} holds with probability $1 - \delta - Cn\exp(-cn\varepsilon^d)$.
+
+	\textcolor{red}{(TODO): Ryan, you asked about AIC, but I'm not clear about whether you want me to (a) make some remark about it, (b) actually go through the work of giving guarantees, or (c) were just curious.}
+
+	\item The upper bound given in equation~\eqref{eqn:laplacian_eigenmaps_estimation_fo} holds with probability $1 - \delta - Cn\exp(-cn\varepsilon^d)$.
 Under the stronger assumption that $f_0 \in C^1(\mc{X};M)$ we can replace the factor of $\delta$ by the sharper $\delta^2/n$, which is less than $\delta$ because $\delta \in (0,1)$.
+Then a routine calculation, which we carry out in Appendix~\ref{subsec:upper_bound_l2pn_risk}, shows that the expected $L^2(P_n)$ loss is on the same order as~\eqref{eqn:laplacian_eigenmaps_estimation_fo}, matching the minimax rate.
 \end{itemize}
@@
 
-\begin{enumerate}[label=(A\arabic*)]
-	\setcounter{enumi}{2}
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{1}
 	\item 
@@
 	\label{eqn:eigenvector_testing_fo}
-	K \asymp \min\Bigl\{(M^2n)^{2d/(4 + d)} \vee 1, n\Bigr\}.
+	K = \min\Bigl\{\floor{(M^2n)^{2d/(4 + d)}} \vee 1, n\Bigr\}.
 	\end{equation}
 \end{enumerate}
-In the following, let $p(d) = \min\{1,2/d\}$.
 \begin{theorem}
@@
 Suppose additionally $f_0 \in H^1(\mc{X},M)$, and that $d < 4$.
-Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_fo}, and if $f_0$ satisfies
+Then there exist constants $C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_fo}, and if $f_0$ satisfies
 	\begin{equation}
 	\label{eqn:laplacian_eigenmaps_testing_criticalradius_fo}
-	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-4/(4 + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{p(d)}} \biggr) \vee \frac{1}{n}
+	\|f_0\|_P^2 \geq C\biggl(\Bigl(M^2(M^2n)^{-4/(4 + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{2/d}} \biggr) \vee \frac{1}{n},
 	\end{equation}
@@
 \end{theorem}
-Although~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} involves taking the maximum of several different terms, the important takeaway of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} is that if $n^{-1/2} \lesssim M^2 \ll n^{1/4}$ ($d = 1$) or $n^{-1/2} \lesssim M^2 \ll n^{2/d - 1/2}$ ($d = 2$ or $d = 3$), then $\varphi$ is a minimax rate-optimal test over $H^1(\mc{X},M)$.
+Although~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} involves taking the maximum of several different terms, the important takeaway of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} is that if $n^{-1/2} \lesssim M \lesssim n^{1/d}$, then $\varphi$ has small worst-case risk as long as $f_0$ is separated from $0$ by at least $M^2(M^2n)^{-4/(4 + d)}$.
+Note that unlike in the estimation setting---where we measured loss in $L^2(P_n)$ error---the separation in~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} is measured in $L^2(P)$ norm.
+Thus~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} implies that $\varphi$ is a minimax rate-optimal test over $H^1(\mc{X};M)$, in the usual sense.
 
@@
 \begin{itemize}
-	\item When $d \geq 4$, as we have already mentioned $H^1(\mc{X})$ does not continuously embed into $\Leb^4(\mc{X})$.
-This has severe consequences for test statistics based on the $L^2(P_n)$-norm---such as $\wh{T}$---or $L^2(P)$ norm---such as $\wt{T}$---since they may no longer have a finite variance.
-Indeed, as far we know when $d \geq 4$ the optimal testing rates over $H^1(\mc{X})$ are not known.
-On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then we show in Appendix~\textcolor{red}{(?)} that the Laplacian eigenmaps test, with $K = n$, has small type II error whenever $\|f_0\|_P^2 \gtrsim n^{-1/2}$.
-	\item The requirement in Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} that $\|f_0\|_P^2 \gtrsim n^{-p(d)}$ is necessary in order for us to establish that with probability at least $1 - 5\delta$,
-	\begin{equation}
-	\label{eqn:small_ball_estimate}
-	\|f_0\|_n^2 \geq \frac{1}{\delta}\|f_0\|_P^2.
-	\end{equation}
-	This requirement prevents us from showing that $\varphi$ is a minimax optimal test when $M \gtrsim n^{1/4}$ ($d = 1$) or $M \gtrsim n^{2/d - 1/2}$ ($d = 2$ or $d = 3$).
-Of course, note that we do show $\varphi$ is an optimal test when $M \asymp 1$, which is by far the most commonly considered case.
-On the other hand, if we assume $f_0 \in C^1(\mc{X};M)$, then we can establish~\eqref{eqn:small_ball_estimate}, and indeed stronger estimates, without any restrictions on $\|f_0\|_P^2$.
-As a result, under this stronger assumption we can show that $\varphi$ is an optimal test for all $n^{-1/2} \lesssim M^2 \lesssim n^{1/d}$, matching our results in the estimation case.
+	\item As mentioned previously, when $d \geq 4$ the first order Sobolev space $H^1(\mc{X})$ does not continuously embed into $\Leb^4(\mc{X})$, and we do not know the optimal rates for regression testing over $H^1(\mc{X},M)$.
+On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then the Laplacian eigenmaps test with $K = n$, has small type II error whenever $\|f_0\|_P^2 \gtrsim Mn^{-1/2}$.
+Note that when $K = n$ the Laplacian eigenmaps test statistic is nothing but the squared $L^2(P_n)$ norm of ${\bf Y}$, $\wh{T} = \|{\bf Y}\|_n^2$.
+See \textcolor{blue}{(Green et al.
+2021)} for details.
+	\item As in the estimation setting, the range of Sobolev ball radii $n^{-1/2} \lesssim M \lesssim n^{1/d}$ for which Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} implies that $\varphi$ is a rate-optimal test covers all those cases for which the critical radius $\epsilon(H^1(\mc{X};M))$ is both $\Omega(1/n)$ and $O(1)$.
 \end{itemize}
@@
 \label{sec:higher_order_sobolev_classes}
-We now assume that the regression function displays some higher-order regularity and is zero-trace, i.e.
+We now consider the situation where the regression function displays some higher-order regularity.
+For reasons already discussed, we also assume the zero-trace boundary condition, i.e.
 $f_0 \in H_0^s(\mc{X})$.
-At first glance, it is not at all clear that Laplacian eigenmaps \emph{can} take advantage of this additional regularity, and it is especially unclear whether it can do so in an optimal manner.
-As $s \to \infty$ and $d$ remains fixed, the optimal estimation and testing rates approach the parametric rate $n^{-1}$.
-On the other hand, the sharpest known results show that Laplacian eigenvectors $v_k$ converge to continuum eigenfunctions $\psi_k$---and thus, for a fixed value of $k$, $\wh{f}$ converges to $\wt{f}$---only at a rate of \textcolor{red}{(?)}, which is obviously much slower than $n^{-1}$.
-
-Nevertheless, as we will establish momentarily, remarkably Laplacian eigenmaps methods continue to be optimal for all orders of $s$, as long as the design density is itself also sufficiently regular, $p \in C^{s - 1}(\mc{X})$.
+We show that Laplacian eigenmaps methods continue to be optimal for all orders of $s$, as long as the design density is itself also sufficiently regular, $p \in C^{s - 1}(\mc{X})$.
 In estimation, this is the case for any dimension $d$, whereas in testing it is the case only when $d \leq 4$.
@@
 In order to show that $\wh{f}$ is an optimal estimator over $H_0^s(\mc{X};M)$, we will require that $\varepsilon$ be meaningfully larger than the lower bound in~\ref{asmp:parameters_estimation_fo}.
-\begin{enumerate}[label=(A\arabic*)]
-	\setcounter{enumi}{3}
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{2}
 	\item 
 	\label{asmp:parameters_estimation_ho}
-	The graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy
+	For constants $c_0$ and $C_0$, the graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy
 	\begin{equation}
 	\label{eqn:radius_ho}
-	\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/d}, (M^2n)^{-1/(2(s - 1) + d)}\biggr\} \ll \varepsilon \ll \min\{1, K^{-1/d}\}
+	C_0\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/d}, (M^2n)^{-1/(2(s - 1) + d)}\biggr\} \leq \varepsilon \leq c_0\min\{1, K^{-1/d}\}
 	\end{equation}
@@
 	\begin{equation*}
-	K \asymp (M^2n)^{2d/(2s + d)}.
+	K = \min\Bigl\{\floor{(M^2n)^{d/(2s + d)}} \vee 1,n\Bigr\}
 	\end{equation*}
 \end{enumerate}
-It is important to note that the two restrictions of~\ref{asmp:parameters_estimation_ho} are not mutually exclusive.
-So long as $M^2 \gg n^{-1}$, then $(M^2n)^{-2/(2(s - 1) + d)} \ll (M^2n)^{-2/(2s + d)}$ for all values of $s$ and $d$.
+Crucially, when $n$ is sufficiently large the two conditions in~\ref{asmp:parameters_estimation_ho} are guaranteed to not be mutually exclusive.
+This is because so long as $M^2 = \omega(n^{-1})$ then $(M^2n)^{-2/(2(s - 1) + d)} = o((M^2n)^{-2/(2s + d)})$, regardless of $s$ and $d$.
 \begin{theorem}
@@
 	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H_0^s(\mc{X},M)$ and $p \in C^{s - 1}(\mc{X})$.
-There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_ho}, then
+There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_ho}, then
 	\begin{equation}
@@
 	\end{equation}
-	with probability at least $1 - \delta - Cn\exp(-cn\varepsilon^d)$.
+	with probability at least $1 - \delta - Cn\exp(-cn\varepsilon^d) - \exp(-K)$.
 \end{theorem}
-Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that in the flat Euclidean setting Laplacian eigenmaps is an in-sample minimax rate optimal estimator over Sobolev classes, for all values of $s$ and $d$.
+Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that in the flat Euclidean setting Laplacian eigenmaps is an in-sample minimax rate-optimal estimator over Sobolev classes, for all values of $s$ and $d$.
 Some other remarks:
@@
 Likewise, certain estimators are ``off the table'', most notably RKHS-based methods such as thin-plate splines of degree $k \leq d/2$.
-Nevertheless, for random design regression with error measured in $\Leb^2(P)$-norm, the spectral series estimator $\wt{f}$ defined in~\eqref{eqn:spectral_series_estimator} obtains the ``usual'' minimax rates $n^{-2s/(2s + d)}$ for all values of $s$ and $d$.
+Nevertheless, for random design regression with error measured in $\Leb^2(P)$-norm, the spectral projection estimator $\wt{f}$ defined in~\eqref{eqn:spectral_series_estimator} obtains the ``usual'' minimax rates $n^{-2s/(2s + d)}$ for all values of $s$ and $d$.
 Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and \ref{thm:laplacian_eigenmaps_estimation_ho} show that the same is true with respect to Laplacian eigenmaps, with error measured in $\Leb^2(P_n)$-norm.
-	\item \textcolor{red}{(TODO 1): This entire comment is too vague right now.
-I'll leave it in, in case Siva + Ryan think it is sufficiently valuable to edit and make more precise.} The lower bound on $\varepsilon$ means that the graph $G$ will be quite dense, with average degree growing polynomially in $n$ as $n \to \infty$.
-Consequently, exactly computing the eigenvectors $v_k$ is more computationally intensive than if the graph radius were on the order of the connectivity threshold $\varepsilon \asymp (\log n/n)^{1/d}$.
-Recall that in the first-order case, we established in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that as long as $\varepsilon \gg (\log n/n)^{1/d}$---meaning average degree growing like $\omega(\log n)$---the estimator $\wh{f}$ could still be optimal.
-Thus Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and~\ref{thm:laplacian_eigenmaps_estimation_ho} can be viewed as revealing a tradeoff between statistical and computational efficiency; although to be clear, we have no theoretical evidence that Laplacian eigenmaps \emph{fails} to adapt to higher-order smoothness when $\varepsilon \asymp (\log n/n)^{1/d}$---we simply cannot prove that it succeeds.
-	\textcolor{red}{(TODO 2): If we decide it is worth our time to investigate the optimal choice of graph radius empirically, we can add a line here mentioning this.}
-
-	Suppose one does choose $\varepsilon$ meaningfully larger than the connectivity threshold, as our theory requires when $s > 1$.
-We now discuss a procedure to efficiently compute an approximation to the Laplacian eigenmaps estimate, without meaningfully increasing the statistical error incurred: \emph{edge sparsification}.
-For a given $\delta \in (0,1)$, various algorithms have been designed \textcolor{red}{(references)} to efficiently remove many edges from the graph $G$, resulting in a sparser graph $\wc{G}$ with Laplacian $\wc{L}_{n,\varepsilon}$, while only perturbing the spectrum of the Laplacian by a factor of at most $\sigma \geq 1$, meaning
-	\begin{equation*}
-	\frac{1}{\sigma} \cdot u^{\top} \wc{L}_{n,\varepsilon} u \leq u^{\top} L_{n,\varepsilon} u \leq \sigma \cdot u^{\top} \wc{L}_{n,\varepsilon}u \quad \textrm{for all $u \in \Reals^n$.}
-	\end{equation*}
-	The graph $\wc{G}$ is said to be $\sigma$-spectrally similar to $G$.
-Let $\wc{f}$ be the Laplacian eigenmaps estimator computed using the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ .
-Because $\wc{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wc{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$, and consequently much faster to compute $\wc{f}$ than $\wh{f}$ \textcolor{red}{(reference needed)}.
-Statistically speaking, as we show in Appendix~\textcolor{red}{(?)} $\wc{f}$ has $L^2(P_n)$-error of at most $\sigma^{2s}$ times the error of $\wh{f}$.
-Therefore, as long as $\sigma \asymp 1$ the estimator $\wc{f}$ will also be rate-optimal.
-Finally, geometric graphs such as $G$ exhibit a special structure; very roughly speaking, no one edge is a bottleneck in geometric graphs.
-As pointed out by~\citet{sadhanala16b}, in this case there exist far simpler and faster methods for sparsification than those of~\textcolor{red}{(references)}; at least empirically, these methods appear to do the job.
 	\item The requirement $p \in C^{s - 1}(\mc{X})$ is a strong condition, but is essential to showing that $\wh{f}$ enjoys faster rates of convergence when $s > 1$.
-We give more detail about this when we discuss our analysis in Section~\ref{subsec:analysis}.
+We explain why in Section~\ref{subsec:analysis}, where we discuss our analysis.
 \end{itemize}
@@
 \paragraph{Testing.} The test $\varphi$ can adapt to the higher-order smoothness of $f_0$, when $\varepsilon$ and $K$ are chosen correctly.
-\begin{enumerate}[label=(A\arabic*)]
-	\setcounter{enumi}{4}
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{3}
 	\item 
@@
 	\label{eqn:eigenvector_testing_ho}
-	K \asymp \min\Bigl\{(M^2n)^{2d/(4s + d)} \vee 1, n\Bigr\}.
+	K = \min\Bigl\{\floor{(M^2n)^{2d/(4 + d)}} \vee 1, n\Bigr\}.
 	\end{equation}
 \end{enumerate}
-When $d \leq 4$, for any value of $s \in \mathbb{N}$ it is possible to choose $\varepsilon$ and $K$ such that both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho} are satisfied, and our next theorem establishes that in this situation $\varphi$ is an optimal test.
-In what follows, let $p(d) = \min\{1,2s/d\}$.
+When $d \leq 4$, for any value of $s \in \mathbb{N}$ when $n$ is sufficiently large it is possible to choose $\varepsilon$ and $K$ such that both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho} are satisfied, and our next theorem establishes that in this situation $\varphi$ is an optimal test.
 \begin{theorem}
@@
 Suppose additionally $f_0 \in H_0^s(\mc{X},M)$, that $p \in C^{s-1}(\mc{X})$, and that $d \leq 4$.
-Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_ho}, and if $f_0$ satisfies
+Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n \geq N$: if the Laplacian eigenmaps test $\varphi$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_ho}, and if $f_0$ satisfies
 	\begin{equation}
 	\label{eqn:laplacian_eigenmaps_testing_criticalradius_ho}
-	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-4s/(4s + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{p(d)}} \biggr) \vee \frac{1}{n},
+	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-4s/(4s + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{2s/d}} \biggr) \vee \frac{1}{n},
 	\end{equation}
@@
 \end{theorem}
-Similarly to the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that $\varphi$ is a minimax optimal test over $H_0^s(\mc{X})$ when $n^{-1/2} \ll M^2 \ll n^{1/4}$ (if $2s < d$), or if $n^{-1/2} \ll M^2 \ll (2/d - 1/2)$ (if $2s \geq d$.)
-
-However, unlike the first-order case, when $4 < d < 4s$ the minimax testing rate over $H_0^s(\mc{X})$ is still $n^{-4s/(4s + d)}$.
+Similarly to the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that $\varphi$ is a minimax optimal test over $H_0^s(\mc{X})$ when $n^{-1/2} \lesssim M^2 \lesssim n^{1/d}$.
+However, unlike the first-order case, when $4 < d < 4s$ the minimax testing rate over $H_0^s(\mc{X})$ is still on the order of $M^2(M^2n)^{-4s/(4s + d)}$.
 Unfortunately, we can no longer claim that $\varphi$ is an optimal test in this regime.
-There do not exist any choices of $\varepsilon$ and $K$ which satisfy both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho}, and as a result we cannot optimally balance testing bias and variance (defined momentarily in~\eqref{eqn:testing_biasvariance}).
-It follows from Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} that, choosing $K \asymp (n)^{-2s/(2s + d)}$ as is optimal for estimation, $\varphi$ must have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2s + d)}$.
-In fact, we can derive slightly better (though still not optimal) rates.
 \begin{theorem}
 	\label{thm:laplacian_eigenmaps_testing_ho_suboptimal}
-	Under the same setup as Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, but with $d > 4$.
-If the Laplacian eigenmaps test $\varphi$ is computed kernel $\eta$ satisfying~\eqref{asmp:kernel}, $K$ satisfying~\eqref{eqn:eigenvector_testing_ho}, and $\varepsilon \asymp K^{-1/d}$, and if
+	Under the same setup as Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, but with $4 < d < 4s$.
+If the Laplacian eigenmaps test $\varphi$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, number of eigenvectors $K$ satisfying~\eqref{eqn:eigenvector_testing_ho}, and $\varepsilon = (M^2n)^{-1/(2(s - 1) + d)}$, and if
 	\begin{equation}
 	\label{eqn:laplacian_eigenmaps_testing_criticalradius_ho_suboptimal}
-	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-2s/(2(s - 1) + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{p(d)}} \biggr) \vee \frac{1}{n},
+	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-2s/(2(s - 1) + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{2s/d}} \biggr) \vee \frac{1}{n},
 	\end{equation}
@@
 \end{theorem}
-We do not know whether the problem here is the test itself, our some looseness in our theoretical bounds, and leave the matter for future work.
-
-\textcolor{red}{(TODO): This last bit is too vague and speculative. Maybe worth deleting entirely?}
-
-A last point: as mentioned at the beginning of this section, the estimation and testing rates we obtain in Theorems~\ref{thm:laplacian_eigenmaps_estimation_ho}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} are superior to those implied by the concentration of graph Laplacian eigenvectors $v_k$ around eigenfunctions $\psi_k$ of the weighted Laplace-Beltrami operator $\Delta_P$.
-This implies (at least) one out of the two following explanations.
-\begin{enumerate}
-	\item The convergence rates of graph eigenvectors to continuum eigenfunctions, $v_k \to \psi_k$, are faster than currently known,  particularly if $p$ exhibits higher-order smoothness.
-	\item The approximation properties of $v_1,\ldots,v_K$ are much stronger than is implied by the concentration of each individual eigenvector $v_k$ around its continuum counterpart $\psi_k$.
-This is reminiscent of similar phenomena exhibited by kernel machines \textcolor{red}{(Belkin 2019)}.
-\end{enumerate}
-Identifying which explanation is responsible would, in our view, be a useful step towards understanding the success of graph Laplacian methods.
+Note that as a consequence of Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, if we choose $K = n^{-2s/(2s + d)}$ then $\varphi$ must have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2s + d)}$.
+Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} shows that $\varphi$ can achieve better, but still not optimal, rates.
+As a technical matter, the problem is that when $d > 4$ there do not exist any choices of $\varepsilon$ and $K$ which satisfy both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho}, and as a result we cannot optimally balance (our upper bound on) testing bias and variance (defined momentarily in~\eqref{eqn:testing_biasvariance}).
+Although we suspect $\varphi$ is truly suboptimal when $d > 4$, technically speaking~\eqref{eqn:testing_biasvariance} gives only an upper bound on testing bias, and thus we cannot rule out that the test $\varphi$ is optimal for all $4 < d < 4s$.
+We leave the matter to future work.
+
+That being said, it is somewhat remarkable that Laplacian eigenmaps \emph{can} take advantage of higher-order smoothness, and especially surprising that it can do so in an optimal manner.
+The sharpest known results show that the graph Laplacian eigenvectors $v_k$ converge to eigenfunctions $\psi_k$ at a rate of \textcolor{red}{(?)}.
+Naively applying these results, one can show that $\wh{f}$ to $\wt{f}$, but only at a rate far slower than the optimal rates for regression.
+Of course when the index $K$ increases with $n$, as is necessary to optimally balance bias and variance, the issue only gets worse.
+Clearly, as a method for regression, the rate of convergence of Laplacian eigenmaps is much better than the rate implied by (what is currently known about) the concentration of individual eigenvectors around their continuum limits.
 
@@
 We now outline the high-level strategy we follow when proving each of Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}.
-We analyze the estimation error of $\wh{f}$, and the testing error of $\wh{\varphi}$, by first conditioning on the design points ${\bf X}$ and deriving \emph{design-dependent} bias and variance terms.
+We analyze the estimation error of $\wh{f}$, and the testing error of $\wh{\varphi}$, by first conditioning on the design points $X_1,\ldots,X_n$ and deriving \emph{design-dependent} bias and variance terms.
 For estimation, we have that with probability at least $1 - \exp(-K)$,
@@
 \paragraph{Estimates on graph quadratic forms.}
-We begin by restating in Proposition~\ref{prop:graph_seminorm_fo} an upper bound on the first-order quadratic form $\dotp{L_{n,\varepsilon}f_0}{f_0}_n$ from \textcolor{blue}{Green et al. 2021}.
+In Proposition~\ref{prop:graph_seminorm_fo} we restate an upper bound on the Dirichlet energy $\dotp{L_{n,\varepsilon}f}{f}_n$ from \textcolor{blue}{Green et al. 2021}. 
 \begin{proposition}[Lemma~1 of \textcolor{blue}{Green et al. 2021}]
 	\label{prop:graph_seminorm_fo}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H^1(\mc{X},M)$.
-There exist constants $c,C$ that do not depend on $f_0$ such that the following statement holds for any $\delta \in (0,1)$: if $\eta$ satisfies~\ref{asmp:kernel} and $\varepsilon < c$, then
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f \in H^1(\mc{X})$.
+There exist constants $c,C$ that do not depend on $f$ or $n$ such that the following statement holds for any $\delta \in (0,1)$: if $\eta$ satisfies~\ref{asmp:kernel} and $\varepsilon < c$, then
 	\begin{equation}
 	\label{eqn:graph_seminorm_fo}
-	\dotp{L_{n,\varepsilon}f_0}{f_0}_n \leq \frac{C}{\delta} M^2,
+	\dotp{L_{n,\varepsilon}f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^1(\mc{X})}^2,
 	\end{equation}
@@
 \end{proposition}
-
-In this work, we establish that a similar result holds for the higher-order case $s > 1$.
-
+Proposition~\ref{prop:graph_seminorm_fo} follows by upper bounding the expectation of $\dotp{L_{n,\varepsilon}f}{f}_n$, which is the Dirichlet energy  $E_{P,\varepsilon}(f) := \dotp{L_{P,\varepsilon}f}{f}_P$, by (a constant times) the squared Sobolev norm $\|f\|_{H^1(\mc{X})}^2$.
+
+In this work, we establish that an analogous bound holds for $\dotp{L_{n,\varepsilon}^sf_0}{f_0}_n$ when $s > 1$. We call this quantity the order-$s$ \emph{graph Sobolev semi-norm}.
 \begin{proposition}
 	\label{prop:graph_seminorm_ho} 
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f_0 \in H_0^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$.
-Then there exist constants $c$ and $C$ that do not depend on $f_0$ such that the following statement holds for any $\delta \in (0,1)$: if $\eta$ satisfies~\ref{asmp:kernel} and $Cn^{-1/(2(s - 1) + d)} < \varepsilon < c$, then
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f \in H_0^s(\mc{X})$ and $p \in C^{s - 1}(\mc{X})$.
+Then there exist constants $c$ and $C$ that do not depend on $f_0$ or $n$ such that the following statement holds for any $\delta \in (0,1)$: if $\eta$ satisfies~\ref{asmp:kernel_flat_euclidean} and $Cn^{-1/(2(s - 1) + d)} < \varepsilon < c$, then
 	\begin{equation}
 	\label{eqn:graph_seminorm_ho}
-	\dotp{L_{n,\varepsilon}^s f}{f}_n \leq \frac{C}{\delta} M^2 \|p\|_{C^{s}(\mc{X})}^2,
+	\dotp{L_{n,\varepsilon}^s f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^s(\mc{X})}^2 ,
 	\end{equation}
@@
 \end{proposition}
-The proof of Proposition~\ref{prop:graph_seminorm_ho} can be divided into two steps.
-In the first step, we show that when $Cn^{-1/(2(s - 1) + d)}$, then with probability at least $1 - \delta$ the graph semi-norm $\dotp{L^s f}{f}_n$ is upper bounded by $1/\delta$ times a non-local continuum energy $E_{\varepsilon}^{(s)}(f)$, defined as
-\begin{equation*}
-E_{\varepsilon}^{(s)}(f) := \dotp{L_{P,\varepsilon}^sf}{f}_{P}.
-\end{equation*}
-Here $L_{P,\varepsilon}$ is a non-local approximation to $\Delta_P$, 
+We now summarize the techniques used to prove Proposition~\ref{prop:graph_seminorm_ho}, which will help explain what role the conditions on $f_0$,$p$ and $\varepsilon$ play.
+To upper bound $\dotp{L_{n,\varepsilon}^sf}{f}_n$ in terms of $\|f\|_{H^s(\mc{X})}^2$, we introduce an intermediate quantity: the \emph{order-s non-local Sobolev seminorm} $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$.
+This seminorm is defined with respect to $L_{P,\varepsilon}$, which is a non-local approximation to $\Delta_P$,
 \begin{equation}
@@
 \end{equation}
-In the second step, we show that whenever $\varepsilon$  is smaller than some contanst $c$, then $E_{\varepsilon}^{(s)}(f)$ is in turn upper bounded by a constant times $M^2 \|p\|_{C^{s - 1}(P)}^2$.
-Note that neither quantity here is random; this is a purely approximation-theoretic statement.
+Then the proof of Proposition~\ref{prop:graph_seminorm_ho} proceeds according to the following steps.
+\begin{itemize}
+	\item First we note that $\dotp{L_{n,\varepsilon}^s f}{f}_n$ is itself a biased estimate of the non-local seminorm $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$.
+Specifically, $\dotp{L_{n,\varepsilon}^s f}{f}_n$ is a $V$-statistic, meaning it is the sum of an unbiased estimator of $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$ (in other words, a $U$-statistic) plus some higher-order, pure bias terms.
+We show that these pure bias terms are negligible when $\varepsilon = \omega(n^{-1/(2(s - 1) + d)})$.
+	\item For $x$ sufficiently in the interior of $\mc{X}$, we show that $L_{P,\varepsilon}^jf(x) \to \sigma_{\eta}^j \Delta_P^jf(x)$ as $\varepsilon \to 0$.
+Here $j = (s - 1)/2$ when $s$ is odd and $j = (s - 2)/2$ when $s$ is even.
+This step bears some resemblance to the analysis of the bias term in kernel smoothing, and requires that $p \in C^{s-1}(\mc{X})$.
+	\item On the other hand for $x$ sufficiently near the exterior of $\mc{X}$, $L_{P,\varepsilon}^jf(x)$ does not converge to $\Delta_P^jf(x)$.
+Instead, we use the zero-trace property of $f$ to show that $L_{P,\varepsilon}^jf(x)$ is small.
+	\item Finally, we combine the results of previous two steps to deduce an upper bound on $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$ in terms of the squared Sobolev norm $\|f\|_{H^s(\mc{X})}^2$.
+Roughly speaking, when $s$ is odd, $\dotp{L_{P,\varepsilon}^sf}{f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}^jf) \approx \sigma_{\eta}^{2j}E_{P,\varepsilon}(\Delta_P^jf)$, whereas when $s$ is even $\dotp{L_{P,\varepsilon}^sf}{f}_P = \|L_{P,\varepsilon}L_{P,\varepsilon}^{j}f\|_{P}^2 \approx \sigma_{\eta}^{2j}\|L_{P,\varepsilon} \Delta_Pf\|_P^2$.
+Reasoning in this way, we can translate estimates of $L_{P,\varepsilon}^jf$ into an upper bound on the order-$s$ non-local Sobolev seminorm, even though $s > j$.
+\end{itemize}
 Together, these steps establish Proposition~\ref{prop:graph_seminorm_ho}.
-
-There are several reasons why this analysis is substantially trickier than the first-order case.
-First of all, we note that when $s > 1$ the quadratic form $\dotp{L_{n,\varepsilon}^s f}{f}_n$ is itself a biased estimate of the non-local energy $E_{\varepsilon}^{(s)}(f)$, in contrast to the first order case where $\Ebb\bigl[\dotp{L_{n,\varepsilon} f}{f}_n\bigr] = E_{\varepsilon}^{(1)}(f)$.
-One can view $\dotp{L_{n,\varepsilon}^s f}{f}_n$ as a $V$-statistic, meaning it is the sum of an unbiased estimator of $E_{\varepsilon}^{(s)}(f)$ (in other words, a $U$-statistic) plus some higher-order, pure bias terms.
-These pure bias terms are negligible only when $\varepsilon \gg n^{-1/(2r + d)}$.
-
-On the other hand, in order to give an estimate of $E_{\varepsilon}^{(s)}(f)$ in terms of the Sobolev norm of $f_0$ and H\"{o}lder norm of $p$, we first show that $L_{P,\varepsilon}^jf(x) \to \Delta_P^jf(x)$ as $\varepsilon \to 0$; here $j = (s - 1)/2$ when $s$ is odd and $j = (s - 2)/2$ when $s$ is even.
-In words, we show that a $j$th-order iterated difference operator approximates a $2j$-th order differential operator.
-In order to establish this fact, we require that on the one hand that $p \in C^{s - 1}(\mc{X})$, and on the other hand that $f$ be zero-trace.
-Crucially we do not try to show $L_{P,\varepsilon}^sf(x) \to \Delta_{P}^sf(x)$: though this may seem like a natural first step towards establishing $\dotp{L_{P,\varepsilon}^sf}{f}_{P} \to \dotp{\Delta_P^sf}{f}_{P}$, note carefully that $\Delta_P^s$ is an order-$2s$ differential operator, whereas we assume that $f$ has only $s$ bounded derivatives.
+It is worth pointing out that we do not try to show $L_{P,\varepsilon}^sf(x) \to \Delta_{P}^sf(x)$.
+This may seem like a natural first step towards a simple proof that $\dotp{L_{P,\varepsilon}^sf}{f}_{P} \to \dotp{\Delta_P^sf}{f}_{P}$.
+The problem is that $\Delta_P^s$ is an order-$2s$ differential operator, whereas we assume that $f$ has only $s$ bounded derivatives.
+Instead we go for the slightly more complicated approach outlined above.
 
@@
 They provide explicit bounds on the relative error $|\lambda_{k} - \lambda_{k}(\Delta_P)|/\lambda_{k}(\Delta_P)$, which show that the relative error is small for sufficiently large $n$ and small $\varepsilon$.
-Crucially, the guarantees hold simultaneously for all $1 \leq k \leq k_{\max}$, where $k_{\max}$ must satisfy $\lambda_{k_{\max}}(\Delta_P) \ll \varepsilon^{-2}$.
-These results are actually stronger than are necessary to establish Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho}---in order to get rate-optimality, we need only show that $\lambda_{K}/\lambda_K(P)$ is $\Omega_P(1)$---but unfortunately they all assume $P$ is supported on a manifold without boundary (i.e.
+Crucially, the guarantees hold simultaneously for all $1 \leq k \leq K$ as long as $\lambda_{K}(\Delta_P) = O(\varepsilon^{-2})$.
+These results are actually stronger than are necessary to establish Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho}---in order to get rate-optimality, we need only show that for the relevant values of $K$, $\lambda_{K}/\lambda_K(P) = \Omega_P(1)$---but unfortunately they all assume $P$ is supported on a manifold without boundary (i.e.
 they assume Model~\ref{def:model_manifold} rather than Model~\ref{def:model_flat_euclidean}).
@@
 	Suppose Model~\ref{def:model_flat_euclidean}.
-Then there exist constants $c$ and $C$ such that the following statement holds: if $\eta$ satisfies~\ref{asmp:kernel} and $C(\log n/n)^{1/d} < \varepsilon < c$, then
+Then there exist constants $c$ and $C$ such that the following statement holds: if $\eta$ satisfies~\ref{asmp:kernel_flat_euclidean} and $C(\log n/n)^{1/d} < \varepsilon < c$, then
 	\begin{equation}
@@
 	\end{equation}
-	with probability at least $1 - Cn\exp\{-n\varepsilon^d\}$. 
+	with probability at least $1 - Cn\exp\{-c n\varepsilon^d\}$. 
 \end{proposition}
 Note immediately that $\lambda_0(\Delta_P) = \lambda_0 = 0$.
-Furthermore, Weyl's Law \textcolor{red}{(Dunlop, Wu)} tells us that under Model~\ref{def:model_flat_euclidean}, $k^{2/d} \lesssim \lambda_{k}(\Delta_P) \lesssim k^{2/d}$ for all $k \in \mathbb{N}, k > 1$.
+Furthermore, Weyl's Law (Lemma~\ref{lem:weyl}) tells us that under Model~\ref{def:model_flat_euclidean}, $k^{2/d} \lesssim \lambda_{k}(\Delta_P) \lesssim k^{2/d}$ for all $k \in \mathbb{N}, k > 1$.
 Combining these statements with~\eqref{eqn:graph_eigenvalue}, we conclude that $\lambda_{K} = \Omega_P(K^{2/d})$ so long as $K \lesssim \varepsilon^{-d}$.
@@
 Finally, in order to show that $\varphi$ has small Type II error whenever~$\|f_0\|_P$ is greater than the critical radius given by~\eqref{eqn:sobolev_space_testing_critical_radius}, we require a lower bound on $\|f_0\|_n^2$ in terms of $\|f_0\|_P^2$.
-In Lemma~\ref{lem:empirical_norm_sobolev} we establish that such a bound holds, whenever $\|f_0\|_P$ is sufficiently large.
-When $f_0 \in H_0^s(\mc{X})$ and $s > d/2$ this is a standard estimate \textcolor{red}{(reference)}, but it seems to be less well-known in the subcritical regime $s \leq d/2$.
+In Proposition~\ref{prop:empirical_norm_sobolev} we establish that such a one-sided bound holds, whenever $\|f_0\|_P$ is sufficiently large.
 \begin{proposition}
 	\label{prop:empirical_norm_sobolev}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f_0 \in H^s(\Xset,M)$.
-There exists a constant $C$ that does not depend on $f_0$ such that the following statement holds for all $\delta \in (0,1)$:  if
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f \in H^s(\Xset,M)$ for some $s > d/4$.
+There exist constants $c$ and $C$ that do not depend on $f_0$ or $n$ such that the following statement holds for any $\delta > 0$:  if
 	\begin{equation}
 	\label{eqn:empirical_norm_sobolev_1}
-	\norm{f_0}_{P} \geq \frac{C M}{\delta}\Bigl(n^{-1/2} \vee n^{-s/d}\Bigr)
-	\end{equation}
-	then
+	\norm{f}_{P} \geq C M \biggl(\frac{1}{\delta n}\biggr)^{s/d}
+	\end{equation}
+	then with probability at least $1 - \exp\{-(cn \wedge 1/\delta)\}$.
 	\begin{equation}
 	\label{eqn:empirical_norm_sobolev}
-	\norm{f_0}_n^2 \geq \delta \|f_0\|_P^2
-	\end{equation}
-	with probability at least $1 - 5\delta$.
+	\norm{f}_n^2 \geq \frac{1}{2} \|f_0\|_P^2.
+	\end{equation}
 \end{proposition}
+To prove Proposition~\ref{prop:empirical_norm_sobolev}, we use the Gagliardo-Nirenberg interpolation inequality to control the $4$th moment of $f$ in terms of $\|f\|_P$ and $|f|_{H^s(\mc{X})}$, then invoke a one-sided Bernstein's inequality as in \cite[Section 14.2]{wainwright2019}.
+Note carefully that the statement~\eqref{eqn:empirical_norm_sobolev} is \emph{not} a uniform guarantee over all $f \in H^s(\mc{X};M)$, as such a statement cannot hold in the sub-critical regime ($2s \leq d$).
+Fortunately, a pointwise bound---meaning a bound that holds with high probability for a single $f \in H^s(\mc{X})$---is sufficient for our purposes.
 
 Finally, invoking the bounds of Propositions~\ref{prop:graph_seminorm_fo}-\ref{prop:empirical_norm_sobolev} inside the bias-variance tradeoffs~\eqref{eqn:estimation_biasvariance} and~\eqref{eqn:testing_biasvariance} and then choosing $K$ to balance bias and variance (when possible), leads to the conclusions of Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}.
+
+\subsection{Computational considerations}
+Recall that when $s = 1$, we have shown that Laplacian eigenmaps is optimal when $\varepsilon \asymp (\log n/n)^{1/d}$ is (up to a constant) as small as possible while still ensuring the graph $G$ is connected.
+On the other hand, when $s > 1$, we can show Laplacian eigenmaps is optimal only when $\varepsilon = \omega(n^{-c})$ for some $c < 1/d$.
+For such a choice of $\varepsilon$, the average degree in $G$ will grow polynomially in $n$ as $n \to \infty$, and computing eigenvectors of the Laplacian of a graph will be more computationally intensive than if the graph were sparse \textcolor{red}{(reference)}.
+Thus Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and~\ref{thm:laplacian_eigenmaps_estimation_ho} can be seen as revealing a tradeoff between statistical and computational efficiency; although to be clear, we have no theoretical evidence that Laplacian eigenmaps \emph{fails} to adapt to higher-order smoothness when $\varepsilon \asymp (\log n/n)^{1/d}$---we simply cannot prove that it succeeds.
+\textcolor{red}{(TODO): If we decide it is worth our time to investigate the optimal choice of graph radius empirically, we can add a line here mentioning this.}
+
+Suppose one does choose $\varepsilon$ meaningfully larger than the connectivity threshold, as our theory requires when $s > 1$.
+We now discuss a procedure to efficiently compute an approximation to the Laplacian eigenmaps estimate, without changing the rate of convergence of the resulting estimator: \emph{edge sparsification}.
+By now there exist various methods see (e.g., the seminal papers of \citet{spielman2011,spielman2013,spielman2014}, or the overview by \citet{vishnoi2012} and references therein) to efficiently remove many edges from the graph $G$ while only slightly perturbing the spectrum of the Laplacian.
+Specifically such algorithms take as input a parameter $\sigma \geq 1$, and return a sparser graph $\wt{G}$, $E(\wt{G}) \subseteq E(G)$, with a Laplacian $\wt{L}_{n,\varepsilon}$ satisfying
+\begin{equation*}
+\frac{1}{\sigma} \cdot u^{\top} \wt{L}_{n,\varepsilon} u \leq u^{\top} L_{n,\varepsilon} u \leq \sigma \cdot u^{\top} \wt{L}_{n,\varepsilon}u \quad \textrm{for all $u \in \Reals^n$.}
+\end{equation*}
+Let $\wt{f}$ be the Laplacian eigenmaps estimator computed using the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ .
+Because $\wt{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wt{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$, and consequently much faster to compute $\wt{f}$ than $\wh{f}$ \textcolor{red}{(reference needed)}.
+Statistically speaking, letting $\wt{\lambda}_k$ be the $k$th eigenvalue of $\wc{L}_{n,\varepsilon}$, we have that conditional on $X_1,\ldots,X_n$,
+\begin{equation*}
+\|\wt{f} - f_0\|_n^2 \leq \frac{\dotp{\wt{L}_{n,\varepsilon}^s f_0}{f_0}_n}{\wt{\lambda}_{K + 1}^s} + \frac{5K}{n} \leq \sigma^{2s} \frac{\dotp{\wt{L}_{n,\varepsilon}^s f_0}{f_0}_n}{\wt{\lambda}_{K + 1}^s} + \frac{5K}{n},
+\end{equation*}
+with probability at least $1 - \exp(-K)$.
+Consequently $\wt{f}$ has $L^2(P_n)$-error of at most $\sigma^{2s}$ times our upper bound on the error of $\wh{f}$, and for any choice of $\sigma$ that is constant in $n$ the estimator $\wt{f}$ will also be rate-optimal.
+
+In fact the aforementioned edge sparsification algorithms are overkill for our needs.
+For one thing, they are designed to work when $\sigma$ is much larger than $1$, whereas in order for $\wc{f}$ to be rate-optimal setting $\sigma$ to be any constant greater than $1$, say $\sigma = 2$, is sufficient.
+Additionally, edge sparsification algorithms are traditionally designed to work in the worst-case, where no assumptions are made on the structure of the graph $G$.
+But the geometric graphs we consider in this paper exhibit a special structure, in which very roughly speaking no single edge is a bottleneck.
+As pointed out by~\citet{sadhanala16b}, in this special case there exist far simpler and faster methods for sparsification, which at least empirically seem to do the job.
 
@@
 In this section we consider the manifold setting, where $(X_1,Y_1),\ldots,(X_n,Y_n)$ are observed according to Model~\ref{def:model_manifold}.
-A theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+A theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
 We build on this work by showing that when $f_0 \in H^s(\mc{X})$ and $P$ is supported on a manifold, Laplacian eigenmaps achieve the sharper minimax estimation and testing rates reviewed in Section~\ref{subsec:minimax_rates_sobolev}.
@@
 \subsection{Laplacian eigenmaps error rates under the manifold hypothesis}
-Unlike in the flat-Euclidean case, because in Model~\ref{def:model_manifold} we assume that $\mc{X}$ is boundaryless it is easy to deal with the first-order $(s = 1)$ and higher-order $(s > 1)$ cases all at once.
-A more distinction between the results we obtain in this section and those of the last is that we will establish Laplacian eigenmaps is optimal only when the regression function $f_0 \in H^s(\mc{X};M)$ for $s \leq 3$.
+Unlike in the flat-Euclidean case, since Model~\ref{def:model_manifold} assumes that $\mc{X}$ is boundaryless it is easy to deal with the first-order $(s = 1)$ and higher-order $(s > 1)$ cases all at once.
+A more important distinction between the results of this section and those of Section~\ref{sec:minimax_optimal_laplacian_eigenmaps} is that we will establish Laplacian eigenmaps is optimal only when the regression function $f_0 \in H^s(\mc{X};M)$ for $s \leq 3$.
 Otherwise, this section will proceed in a similar fashion to Section~\ref{sec:higher_order_sobolev_classes}.
@@
 
-\begin{enumerate}[label=(A\arabic*)]
-	\setcounter{enumi}{5}
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{4}
 	\item 
@@
 	\label{asmp:parameters_estimation_manifold}
-	The graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy
+	For a constant $C_0$, the graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy
 	\begin{equation}
 	\label{eqn:radius_estimation_manifold}
-	\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/m}, n^{-1/(2(s - 1) + m)}\biggr\} \ll \varepsilon \ll \min\{1, K^{-1/m}\}
+	C_0\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/m}, n^{-1/(2(s - 1) + m)}\biggr\} \leq \varepsilon \leq \min\{i_0, K^{-1/m}\}
 	\end{equation}
@@
 	\begin{equation*}
-	K \asymp (M^2n)^{2m/(2s + m)}.
+	K = \min\Bigl\{\floor{(M^2n)^{m/(2s + m)}} \wedge 1,n \Bigr\}.
 	\end{equation*}
@@
 	\end{equation}
-	with probability at least $1 - \delta - Cn\exp(-cn\varepsilon^m)$.
+	with probability at least $1 - \delta - Cn\exp(-cn\varepsilon^m) - \exp(-K)$.
 \end{theorem}
@@
 Likewise, to construct a minimax optimal test using $\wh{T}$, we choose $\varepsilon$ and $K$ as in~\ref{asmp:parameters_testing_fo}, except with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
-\begin{enumerate}[label=(A\arabic*)]
-	\setcounter{enumi}{7}
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{5}
 	\item 
@@
 	\begin{equation*}
-	K \asymp C_2 (M^2n)^{2m/(4s + m)}
+	K = \min\Bigl\{\floor{(M^2n)^{2m/(4s + m)}} \wedge 1,n \Bigr\}.
 	\end{equation*}
@@
 Suppose additionally $f_0 \in H^s(\mc{X},M)$, that $p \in C^{s-1}(\mc{X})$, and that $s \leq 3$ and $m \leq 4$.
-Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed kernel $\eta$ satisfying~\ref{asmp:kernel_manifold}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_manifold}, and if $f_0$ satisfies
+Then there exist constants $c$, $C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed kernel $\eta$ satisfying~\ref{asmp:kernel_manifold}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_manifold}, and if $f_0$ satisfies
 	\begin{equation}
 	\label{eqn:laplacian_eigenmaps_testing_criticalradius_manifold}
-	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-4s/(4s + m) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{p(m)}} \biggr) \vee \frac{1}{n},
+	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-4s/(4s + m) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{2s/d}} \biggr) \vee \frac{1}{n},
 	\end{equation}
@@
 	\item Unlike in the full-dimensional case, our upper bounds on the estimation and testing error of Laplacian eigenmaps match the minimax rate only when $s \leq 3$.
-Our upper bounds when $s \geq 4$ follow from the embedding $H^s(\mc{X};M) \subset H^{3}(P;M)$, i.e they match the rates we get just by assuming that the $3$rd order derivative is bounded, and are clearly suboptimal.
+Our upper bounds when $s \geq 4$ follow from the embedding $H^s(\mc{X};M) \subset H^{3}(\mc{X};M)$, i.e they match the rates we get just by assuming that the $3$rd order derivative is bounded, and are clearly suboptimal.
 
@@
 
-	As this explanation suggests, by building $G$ using the geodesic distance one could avoid this error, and would obtain superior rates of convergence.
+	As this explanation suggests, by building $G$ using the geodesic distance one could avoid this error, and might obtain superior rates of convergence.
 However this is not an option for us, as we assume $\mc{X}$---and in particular its geodesics---are unknown.
-Likewise, traditional spectral series estimators achieve the minimax rate for all values of $s$ and $m$, but are undesirable for the same reason---we do not want to assume that $\mc{X}$ is known.
-It is not clear whether this gap between spectral series and Laplacian eigenmaps estimators---or more generally, between estimators which assume the manifold is known, and those which do not---is real, or a product of loose upper bounds.
+Likewise, a classical spectral projection estimator, using eigenfunctions of the manifold Laplace-Beltrami operator, will achieve the minimax rate for all values of $s$ and $m$; but this is undesirable for the same reason---we do not want to assume that $\mc{X}$ is known.
+It is not clear whether this gap between spectral projection and Laplacian eigenmaps estimators---or more generally, between estimators which assume the manifold is known, and those which do not---is real, or a product of loose upper bounds.
 
@@
 
-\textcolor{red}{(TODO)}
-\begin{itemize}
-	\item \textcolor{red}{Give a reference regarding the optimality of spectral series estimators on known manifolds.}
-\end{itemize}
+\paragraph{Analysis.}
+The high-level strategy used to prove Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold} is the same as in the flat-Euclidean setting.
+More specifically, we will use precisely the same bias-variance decompositions~\eqref{eqn:estimation_biasvariance} (for estimation) and~\eqref{eqn:testing_biasvariance} (for testing).
+The difference will be that our bounds on the graph Sobolev seminorm $\dotp{L_{n,\varepsilon}^sf_0}{f_0}_n$, graph eigenvalue $\lambda_K$, and empirical norm $\|f_0\|_n^2$ will now always depend on the intrinsic dimension $m$, rather than the ambient dimension $d$.
+The precise results we use are contained in Propositions~\ref{prop:graph_seminorm_manifold}-\ref{prop:empirical_norm_sobolev_manifold}.
+\begin{proposition}
+	\label{prop:graph_seminorm_manifold} 
+	Suppose Model~\ref{def:model_manifold}, and additionally that $f_0 \in H^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$ for $s = 1,2$ or $3$.
+Then there exist constants $c_0,C_0$ and $C$ that do not depend on $f_0$, $n$ or $M$ such that the following statement holds for any $\delta \in (0,1)$: if $\eta$ satisfies~\ref{asmp:kernel_manifold} and $C_0n^{-1/(2(s - 1) + m)} < \varepsilon < c_0$, then
+	\begin{equation}
+	\label{eqn:graph_seminorm_manifold}
+	\dotp{L_{n,\varepsilon}^s f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^s(\mc{X})}^2,
+	\end{equation}
+	with probability at least $1 - 2\delta$.
+\end{proposition}
+
+As discussed previously, when $\mc{X}$ is a domain without boundary and $\Delta_P$ is the manifold weighted Laplace-Beltrami operator, appropriate bounds on the graph eigenvalues $\lambda_k$ have already been derived in \citep{burago2014,trillos2019,garciatrillos19}.
+The precise result we need is a simple consequence of Theorem 2.4 of~\citep{calder2019}.
+\begin{proposition}[\textbf{c.f Theorem 2.4 of~\citep{calder2019}}]
+	\label{prop:graph_eigenvalue_manifold}
+	Suppose Model~\ref{def:model_manifold}.
+Then there exist constants $c$ and $C$ such that the following statement holds: if $\eta$ satisfies~\ref{asmp:kernel_manifold} and $C(\log n/n)^{1/m} < \varepsilon < c$, then
+	\begin{equation}
+	\label{eqn:graph_eigenvalue_manifold}
+	\lambda_k \geq c \cdot \min\Bigl\{\lambda_k(\Delta_P), \frac{1}{\varepsilon^{2}} \Bigr\} \quad \textrm{for all $1 \leq k \leq n$,}
+	\end{equation}
+	with probability at least $1 - Cn\exp\{-c n\varepsilon^d\}$. 
+\end{proposition}
+(For the specific computation used to deduce Proposition~\ref{prop:graph_eigenvalue_manifold} from Theorem 2.4 of~\citep{calder2019}, see~\textcolor{blue}{(Green 2021)}.)
+
+Finally, using a Gagliardo-Nirenberg inequality for functions on compact Riemmanian manifolds, we obtain a lower bound on empirical norm $\|f\|_n$ under the hypotheses of Model~\ref{def:model_manifold}.
+\begin{proposition}
+	\label{prop:empirical_norm_sobolev_manifold}
+	Suppose Model~\ref{def:model_manifold}, and additionally that $f_0 \in H^s(\Xset,M)$ for some $s > m/4$.
+There exists a constant $C$ that does not depend on $f_0$ such that the following statement holds for all $\delta > 0$:  if
+	\begin{equation}
+	\label{eqn:empirical_norm_sobolev_manifold_1}
+	\norm{f_0}_{P} \geq \frac{C M}{\delta^{s/m}}n^{-s/m},
+	\end{equation}
+	then with probability at least $1 - \exp\{-(cn \wedge 1/\delta)\}$,
+	\begin{equation}
+	\label{eqn:empirical_norm_sobolev_manifold}
+	\norm{f_0}_n^2 \geq \frac{1}{2} \|f_0\|_P^2.
+	\end{equation}
+\end{proposition}
 
@@
 \begin{equation*}
-\bigl(T_{h,Q}f)(x) := \frac{1}{d_Q(x)} \int_{\Omega} f(z)\psi\biggl(\frac{\|z - x\|}{h}\biggr) \,dQ(z),
-\end{equation*}
-where $d_Q(x) := \int_{\Omega} \psi\bigl(\|z - x\|/\varepsilon\bigr) \,dQ(z)$.
-For convenience, we will denote $T_{\varepsilon,P_n}$ by $T_{\varepsilon,n}$, and $d_{P_n}$ by $d_n$.
-We extend the Laplacian eigenmaps estimator by passing the kernel smoother $T_{h,n}$ over it, that is we consider the estimator $T_{h,n}\wh{f}$, which is defined at every $x \in \mc{X}$.
+\bigl(T_{Q,h}f)(x) := 
+\begin{dcases*}
+\frac{1}{d_{Q,h}(x)} \int_{\Omega} f(z)\psi\biggl(\frac{\|z - x\|}{h}\biggr) \,dQ(z), & \textrm{if $d_{Q,h}(x) > 0$,} \\
+0, &\textrm{otherwise,}
+\end{dcases*}
+\end{equation*}
+where $d_{Q,h}(x) := \int_{\Omega} \psi\bigl(\|z - x\|/\varepsilon\bigr) \,dQ(z)$.
+For convenience, we will write $T_{\varepsilon,n}f(x) := T_{\varepsilon,P_n}f(x)$, and $d_{n,h}(x) := n \cdot d_{P_n,h}(x)$.
+We extend the Laplacian eigenmaps estimator by passing the kernel smoother $T_{h,n}$ over it, that is we consider the estimator $T_{h,n}\wh{f}$, which is defined at every $x \in \mc{X}$ (indeed, at every $x \in \Rd$).
+Note that ``extension'' here is a slight abuse of nomenclature, since $T_{h,n}\wh{f}(X_i)$ and $\wh{f}_i$ may not agree in-sample.
 
@@
 
-In Lemma~\ref{lem:kernel_smoothing_laplacian_eigenmaps}, we consider an arbitrary estimator $\wc{f} \in L^2(P_n)$.
+In Lemma~\ref{lem:kernel_smoothing_insample}, we consider an arbitrary estimator $\wc{f} \in L^2(P_n)$.
 We show that the out-of-sample error $\|T_{n,h}\wc{f} - f_0\|_P^2$ can be upper bounded by three terms--- (a constant times) the in-sample error $\|\wc{f} - f_0\|_n^2$, and variance and bias terms that arise naturally in the analysis of kernel smoothing over noiseless data..
-We shall assume the following conditions on $\psi$.
+We shall assume the following conditions on $\psi$ and $h$.
 \begin{enumerate}[label=(K\arabic*)]
-	\setcounter{enumi}{1}
+	\setcounter{enumi}{2}
 	\item
@@
 	\begin{equation*}
-	\int_{-\infty}^{\infty} \psi(|s|) \,ds = 1.
+	\int_{-\infty}^{\infty} \psi(|z|) \,dz = 1.
+	\end{equation*}
+\end{enumerate}
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{6}
+	\item
+	\label{asmp:bandwidth}
+	For constants $c_0$ and $C_0$, the bandwidth parameter $h$ satisfies
+	\begin{equation*}
+	C_0\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq h \leq c_0.
 	\end{equation*}
@@
 \begin{lemma}
-	\label{lem:kernel_smoothing_laplacian_eigenmaps}
+	\label{lem:kernel_smoothing_insample}
 	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $\wc{f} \in L^2(P_n)$, $f_0 \in H^1(\mc{X})$ and $p \in C^1(\mc{X})$.
-If the kernel smoothing estimator $T_{h,n}\wc{f}$ is computed with kernel $\psi$ satisfying~\ref{asmp:kernel}, it holds that
-	\begin{equation}
-	\label{eqn:kernel_smoothing_laplacian_eigenmaps}
+If the kernel smoothing estimator $T_{h,n}\wc{f}$ is computed with kernel $\psi$ satisfying~\ref{asmp:kernel} and bandwidth $h$ satisfying~\ref{asmp:bandwidth}, it holds that
+	\begin{equation}
+	\label{eqn:kernel_smoothing_insample}
 	\|T_{n,h}\wc{f} - f_0\|_P^2 \leq C\biggl(\|\wc{f} - f_0\|_n^2 + \frac{1}{\delta} \cdot \frac{h^2}{nh^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{h,P}f_0 - f_0\|_P^2\biggr),
@@
 Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data, by a factor of $h^2$.
-The bias term, on the other hand, is standard.
-A standard analysis shows that the $\|T_{h,P}f_0 - f_0\|_P^2 \lesssim \varepsilon^{2s}$ when $\psi$ is an order-$s$ kernel, meaning
-\begin{equation*}
-\int_{-\infty}^{\infty} \psi(|s|) = 1,~~\int_{-\infty}^{\infty} t^j \psi(|t|) \,dt = 0~~\textrm{for $j = 1,\ldots,s + d - 2$},~~\textrm{and}~~\int_{-\infty}^{\infty} t^{s + d - 1} \psi(|t|) \,dt < \infty.
-\end{equation*}
+On the other hand the bias term is typical.
+When $\psi$ is an order-$s$ kernel, a standard analysis shows that the $\|T_{h,P}f_0 - f_0\|_P^2 \lesssim \varepsilon^{2s}$.
+\begin{enumerate}[label=(K\arabic*)]
+	\setcounter{enumi}{3}
+	\item
+	\label{asmp:ho_kernel}
+	The kernel function $\psi$ is an order-$s$ kernel, meaning that it satisfies
+	\begin{equation*}
+	\int_{-\infty}^{\infty} \psi(|z|) \,dz = 1, \quad \int_{-\infty}^{\infty} z^j \psi(|z|) \,dz = 0 ~~\textrm{for}~j = 1,\ldots, s + d - 2, \quad \textrm{and}~ \int_{-\infty}^{\infty} z^{s + d - 1} \psi(|z|) \,dz < \infty.
+	\end{equation*}
+\end{enumerate}
 
@@
 Will need to double check closely, since I could not find a reference analyzing kernel smoothing using a radially symmetric higher order kernel when $d > 1$.}
-\begin{lemma}
-	\label{lem:kernel_smoothing_bias}
-	Suppose Model~\ref{def:model_flat_euclidean}, and that the smoothing operator $T_{h,P}$ is computed with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel}.
+
+Choosing $h \asymp n^{-1/(2(s - 1) + d)}$ balances the kernel smoothing bias and variance terms in~\eqref{eqn:kernel_smoothing_insample}, and implies that
+\begin{equation}
+\label{eqn:kernel_smoothing_insample2}
+\|T_{n,h}\wc{f} - f_0\|_P^2 \leq C\biggl(\|\wc{f} - f_0\|_n^2 + \frac{1}{\delta}n^{-2s/(2(s - 1) + d)}\biggr).
+\end{equation}
+\eqref{eqn:kernel_smoothing_insample2} tells us that the additional error incurred by passing a kernel smoother over an in-sample estimator $\wc{f}$ is negligible compared to the minimax rate of estimation.
+Consequently, if $\wc{f}$ converges at the minimax rate in $L^2(P_n)$, then $T_{n,h}\wc{f}$ will converge at the minimax rate in $L^2(P)$.
+It follows immediately from Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} (when $s = 1$) or Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} (when $s > 1$), that $T_{h,n}\wh{f}$ achieves the optimal rate of convergence in $L^2(P)$.
+
+\begin{theorem}
+	\label{thm:laplacian_eigenmaps_estimation_out_of_sample}
+	Suppose Model~\ref{def:model_flat_euclidean}.
+There exist constants $c$, $C$, and $N$ that do not depend on $f_0$ or $n$ such that each the following statements hold with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\} - Ch^d\exp\{-cnh^d\}$,  for all $n \geq N$ and for any $\delta \in (0,1)$.
 	\begin{itemize}
-		\item If $f_0 \in H^1(\mc{X})$, it follows that
-		\begin{equation*}
-		\|T_{h,P}f_0 - f_0\|_P^2 \leq C h^{2} |f|_{H^1(\mc{X})}^2.
-		\end{equation*}
-		\item If $f_0 \in H_0^{s}(P)$, $p \in C^{s}(\mc{X})$, and $\psi$ is an order-$s$ kernel, it follows that
-		\begin{equation*}
-		\|T_{h,P}f_0 - f_0\|_P^2 \leq C h^{2s} |f|_{H^s(\mc{X})}^2.
+		\item If $f_0 \in H^1(\mc{X};M)$, the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_fo}, and the out-of-sample extension $T_{h,n}\wh{f}$ is computed with bandwidth $h = n^{-1/d}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel}, then
+		\begin{equation*}
+		\|T_{h,n}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2s/(2s + d)}.
+		\end{equation*}
+		\item If $f_0 \in H_0^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$ for some $s \in \mathbb{N}, s > 1$, and the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_ho}, and the out-of-sample extension $T_{h,n}\wh{f}$ is computed with bandwidth $h = n^{-1/(2(s - 1) + d)}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel} and~\ref{asmp:ho_kernel}, then
+		\begin{equation*}
+		\|T_{h,n}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2s/(2s + d)}.
 		\end{equation*}
 	\end{itemize}
-\end{lemma}
-Choosing $h \asymp n^{-1/(2(s - 1) + d)}$ balances the kernel smoothing bias and variance terms in~\eqref{eqn:kernel_smoothing_laplacian_eigenmaps}.
-Then, applying the bounds in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} (when $s = 1$) or Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} (when $s > 1$), we conclude that $T_{h,n}\wh{f}$ is an optimal estimator, with respect to error measured in $L^2(P)$ norm.
-
-% AG: Get rid of this!
-% \begin{theorem}
-%	\label{thm:laplacian_eigenmaps_estimation_out_of_sample}
-%	Suppose we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to~\eqref{eqn:model}.
-Suppose that for some $s \in \mathbb{N}$, the regression function $f_0 \in H^s(\mc{X};M)$ (if $s = 1$) or $H_0^s(\mc{X};M)$ (if $s > 1$).
-Suppose also that the distribution $P$ has support $\mc{X}$ which satisfies~\ref{asmp:domain} and density $p \in C^s(\mc{X};M)$ which satisfies~\ref{asmp:density}.
-Then if the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_ho}, then the out-of-sample extension $T_{h,n}\wh{f}$ computed with bandwidth $h \asymp n^{-1/(2(s - 1) + d)}$ satisfies
-%	\begin{equation*}
-%	\|T_{h,n}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^{2d/(2s + d)}n^{-2s/(2s + d)},
-%	\end{equation*}
-%	with probability at least $1 - \delta - Cn\exp\{-Cn\varepsilon^d\} - Ch^d\exp\{-Cnh^d\}$. 
-% \end{theorem}
-
-Some other remarks:
+\end{theorem}
+Some remarks:
 \begin{itemize}
-	\item Extending the estimator $\wh{f}$ to be defined out-of-sample allows us to select the hyper-parameters $\varepsilon$ and $K$ using sample splitting or cross validation.
+	\item Since $T_{n,h}\wh{f}$ is defined out-of-sample, we can use sample splitting or cross validation methods to tune hyperparameters, which we could not do for the original estimator $\wh{f}$.
 For instance, we can (i) split the sample into two halves, (ii) use the first half to compute $T_{h,n}\wh{f}$ for various values of $\varepsilon$, $h$, and $K$, (iii) choose the optimal values of these three hyperparameters by minimizing error on the held out set.
-This is by far the most common used way to choose hyperparameters \textcolor{red}{(references)}, and it addresses a serious limitation of Laplacian eigenmaps.
-Moreover, it is known \textcolor{red}{(references)} that choosing hyper-parameters by minimizing error over a held-out set can result in estimators that optimally adapt to the order of regularity $s$---that is, that are rate-optimal (up to $\log$ factors), even when $s$ is unknown.
-We believe that similar arguments will imply that $T_{h,n}\wh{f}$ is adaptive in this sense, though we do not pursue the details.
-\textcolor{red}{(TODO): Siva would like me to pursue the details.}
-	\item We consider extension by kernel smoothing because it is a simple and statistically optimal procedure that does not require any knowledge of the domain $\mc{X}$ or distribution $P$---as we have argued, this latter property is one of the main selling points of Laplacian eigenmaps as a tool for nonparametric regression.
-
-	That being said, there exist many other approaches to extending (or interpolating) a function $f$ using only evaluations $\{f(X_1),\ldots,f(X_n)\}$, and we now comment on a few of these alternatives.
-One approach is to perform minimum norm interpolation, where one defines a normed space $(\mc{F},\|\cdot\|_{\mc{F}})$ and then solves the optimization problem
+Practically speaking, cross-validation is one of the most common approaches to choosing hyperparameters.
+Theoretically, it is known \textcolor{red}{(references)} that choosing hyper-parameters through sample splitting can result in estimators that optimally adapt to the order of regularity $s$.
+In other words, it leads to estimators that are rate-optimal (up to $\log n$ factors), even when $s$ is unknown.
+Similar arguments should imply that $T_{h,n}\wh{f}$ is adaptive in this sense when $\varepsilon,h$ and $K$ are chosen by sample splitting.
+\textcolor{red}{(TODO): Siva would like me to fill in some details?}
+	\item There exist many other approaches to extending a function $f$ using only evaluations $\{f(X_1),\ldots,f(X_n)\}$.
+We consider extension by kernel smoothing because it is a simple and statistically optimal procedure that does not require any knowledge of the domain $\mc{X}$ or distribution $P$---as we have argued, this latter property is one of the main selling points of Laplacian eigenmaps as a tool for nonparametric regression.
+
+	We now comment on a few alternative methods. One such approach is minimum norm interpolation. Here one defines a normed space $(\mc{F},\|\cdot\|_{\mc{F}})$ and then solves the optimization problem
 	\begin{equation*}
@@
 Another, arguably simpler approach is to extend $f$ to be piecewise constant on the Voronoi tessellation induced by $X_1,\ldots,X_n$, or equivalently to perform $1$-nearest neighbors regression on $f$.
-However, this approach is theoretically optimal only when $f_0 \in C^1(\mc{X})$, in contrast to the kernel smoothing method we propose and study.
+However, this approach is theoretically optimal only when $f_0$ is Lipschitz, in contrast to the kernel smoothing method we propose and study.
 
@@
 This is in one sense a strength, since they can be used to extrapolate any estimator defined only in-sample.
-But it also potentially a weakness.
-Each of these methods have their own statistical and approximation errors (bias and variance) which can be fundamentally different than those of Laplacian eigenmaps, and there is a danger that in extrapolating the Laplacian eigenmaps estimator in this way we are taking the ``worst of both worlds''.
+But it is also potentially a weakness.
+Each of these methods have their own approximation and estimation errors (bias and variance) which can be fundamentally different than those of Laplacian eigenmaps, and there is a danger that in extrapolating the Laplacian eigenmaps estimator in this way we are taking the ``worst of both worlds''.
 Our theory shows that when the data model is Model~\ref{def:model_flat_euclidean} and we perform extrapolation by kernel smoothing, this is not a problem, at least in a minimax sense.
@@
 \begin{itemize}
-	\item \textbf{Laplacian eigenmaps approximates spectral series.} $1$d unit cube, uniform density, ``interesting'' regression function.
-Compare Laplacian eigenmaps with spectral series estimator, including their MSE curves, and their fits at several different values of $K$.
+	\item \textbf{Laplacian eigenmaps approximates spectral projection.} $1$d unit cube, uniform density, ``interesting'' regression function.
+Compare Laplacian eigenmaps with spectral projection estimator, including their MSE curves, their fits at several different values of $K$, and their risk as a function of $n$.
 	\item \textbf{Laplacian eigenmaps beats nonparametric least squares.} $1$d unit cube, very non-uniform density, same regression function.
-Make the same three comparisons, but in this case with respect to the least-squares estimator.
-	\item \textbf{Investigate our theoretical bounds on hyperparameters.} $1$d unit cube, uniform density, ``smooth'' regression function.
+Compare risk of Laplacian eigenmaps to nonparametric least squares (with eigenfunctions of unweighted Laplace operator as features) as a function of $n$.
+Show that eigenmaps outperforms least squares.
+	\item \textbf{Investigate our theoretical requirements for hyperparameters.} $1$d unit cube, uniform density, ``smooth'' regression function.
 Investigate the MSE heat map of Laplacian eigenmaps, as a function of $K$ and $\varepsilon$.
+In particular, see if the best combination of these hyperparameters involves a relatively large choice of bandwidth $\varepsilon$.
+Possibly separately plot condtional-on-$X$ bias and variance, as a function of $K$, for a few different choices of $\varepsilon$.
 	\item \textbf{Manifold setting.} Swiss roll, density \textcolor{red}{(?)}, regression function \textcolor{red}{(?)}.
 Evaluate Laplacian eigenmaps in-sample error, and Laplacian-eigenmaps + kernel smoothing out-of-sample error.
+	\item \textbf{Laplacian eigenmaps beats Laplacian smoothing.} High-dimensional unit cube, uniform density, ``interesting'' regression function.
+Compare risk of Laplacian eigenmaps to risk of Laplacian smoothing as a function of $n$, show that the former outperforms the latter.
 \end{itemize}
@@
 \subsection{Comparison with other estimators}
-In this paper, we have motivated Laplacian eigenmaps by viewing it as a noisy approximation of a classical spectral series method, which is its most obvious counterpart.
+In this paper, we have motivated Laplacian eigenmaps by viewing it as a noisy approximation of a classical spectral projection method, which is its most obvious counterpart.
+We have shown that Laplacian eigenmaps inherits the optimality properties of its more classical counterpart.
 We now discuss the relationship between Laplacian eigenmaps and three other approaches to nonparametric regression.
@@
 \paragraph{Nonparametric least squares.}
-The standard recommended alternative to spectral series methods, when the distribution $P$ is considered non-uniform or unknown, is to do least-squares.
-Concretely, suppose instead we had access only to eigenfunctions $\phi_1,\phi_2,\ldots$ of the unweighted Laplace-Beltrami operator $\Delta = \Delta_{\nu}$.
-Then letting $\Phi_K = \mathrm{span}\{f_1,\ldots,f_K\}$, the least-squares estimator and test statistic
+The standard recommended alternative to spectral projection methods, when the distribution $P$ is considered non-uniform or unknown, is to do least-squares.
+For example, suppose instead of knowing $\psi_1,\psi_2,\ldots$, we had access only to eigenfunctions $\phi_1,\phi_2,\ldots$ of an unweighted Laplace-Beltrami operator $\Delta$.
+Then letting $\Phi_K = \mathrm{span}\{\phi_1,\ldots,\phi_K\}$, the least-squares estimator and test statistic
 \begin{equation*}
@@
 are still rate-optimal over $H_0^s(\mc{X})$.
+This holds true for both Model~\ref{def:model_flat_euclidean} and~\ref{def:model_manifold}, and is a consequence of our assumption that $p$ is bounded away from $0$.
 
 However, this is not a totally satisfactory fix.
-For one thing, the least squares approach still requires that we know $\Delta$, and is not well suited for the case where $\mc{X}$ is an unknown manifold, and $\Delta$ is the manifold Laplace-Beltrami operator.
-Additionally, even if we know $\mc{X}$, diagonalizing the Laplace-Beltrami operator $\Delta$ is quite difficult for all but a few special domains, such as $\mc{X} = [0,1]^d$.
+For one thing, the least squares approach just outlined requires that we know the domain $\mc{X}$, in the strong sense that we know the Laplace-Beltrami operator $\Delta$ defined on $\mc{X}$.
+Domain knowledge is generally precious information, and such strong knowledge of $\mc{X}$ seems particularly unrealistic in the case where $\mc{X}$ is a manifold, and $\Delta$ is the manifold Laplace-Beltrami operator.
+Additionally, even if we know $\mc{X}$, diagonalizing the Laplace-Beltrami operator $\Delta$ is quite difficult for all but a few special domains, such as the unit cube $\mc{X} = [0,1]^d$ or torus $\mc{X} = \mathbb{T}^d$.
 
@@
 using the estimator $T_{h,n}Y$ (possibly for a different choice of $h$).
-In Appendix~\textcolor{red}{(?)}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\|f_0 - T_{h,n}\wh{f}\|_P^2 \ll \inf_{h'} \|f_0 - T_{h',n}Y\|_P^2$.
+In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{h,n}\wh{f}\|_P^2$ is of a strictly lower order than $\inf_{h'} \Ebb\|f_0 - T_{h',n}Y\|_P^2$.
 This is possible because Laplacian eigenmaps induces a completely different bias than kernel smoothing.
-In particular, when $f_0$ and $p$ satisfy the so-called \emph{cluster} assumption--- i.e.
-$f_0$ is piecewise constant in high-density regions of $p$--- then the bias of Laplacian eigenmaps can much smaller than that of kernel smoothing (for equivalent levels of variance).
+For example, when $f_0$ and $p$ satisfy the so-called \emph{cluster} assumption--- e.g.
+$f_0$ is piecewise constant in high-density regions (clusters) of $p$--- then the bias of Laplacian eigenmaps can much smaller than that of kernel smoothing (for equivalent levels of variance).
 
@@
 There has been some work \textcolor{red}{(Rigollet, Wasserman, Niyogi, El Alaoui)} analyzing semi-supervised learning under various instantiations of the cluster assumption.
-However, a comprehensive analysis of various methods, including Laplacian eigenmaps, in this setting remains outstanding.
+However, a comprehensive analysis of various methods, including Laplacian eigenmaps and kernel smoothing, in this setting remains outstanding.
 
 \paragraph{Laplacian smoothing.}
-\textcolor{red}{(TODO)}
+As mentioned previously, graph Laplacian smoothing uses the graph Laplacian $L_{n,\varepsilon}$ to form the penalty in a penalized least squares estimator.
+It can be calculated by one solve of a sparse, diagonally dominant linear system.
+Thus it should be much faster to compute than Laplacian eigenmaps, in which one must find (many) eigenvectors of $L_{n,\varepsilon}$.
+On the other hand, Laplacian smoothing is known to be minimax rate-optimal only in very limited regimes (over the Sobolev spaces $H^1(\mc{X})$ for $1 \leq d < 4$.) In contrast, we have shown that Laplacian eigenmaps is minimax rate-optimal for all $d$ (when $s = 1$) .
+We have also shown that Laplacian eigenmaps can adapt to higher-order smoothness, i.e.
+it can be minimax rate-optimal when $s > 1$.
+Thus, the known statistical properties of Laplacian eigenmaps are much stronger than those of Laplacian smoothing.
 
@@
 
-Much is known about classical spectral series methods beyond their rate optimality.
-An incomplete list: such estimators and tests exhibit \emph{sharp optimality}, meaning their risk is within a $(1 + o(1))$ factor of the optimal risk; they can adapt to unknown smoothness of the regression function; they can be used to estimate smooth functionals of the regression function;  finally, they can be used to form confidence sets in $L^2(P)$.
+Much is known about classical spectral projection methods beyond their rate optimality.
+For instance: such estimators and tests exhibit \emph{sharp optimality}, meaning their risk is within a $(1 + o(1))$ factor of the optimal risk; they can adapt to unknown smoothness of the regression function; they can be used to estimate smooth functionals of the regression function;  finally, they can be used to form confidence sets in $L^2(P)$.
 It would be interesting to see if Laplacian eigenmaps could replicate the performance of classical methods in any, or all, of these problems.

### 2021-04-28 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex

@@
 Geometric graphs encode both local information and global information about $P$ in an extremely general manner.
-For this reason they have been leveraged to conduct many different fundamental statistical tasks, such as clustering, manifold learning, semi-supervised learning, classification, and regression \textcolor{red}{(TODO): references}.
+For this reason they have been leveraged to conduct many different fundamental statistical tasks, such clustering, manifold learning, semi-supervised learning, classification, and regression \textcolor{red}{(TODO): references}.
 Substantial theoretical progress has been made regarding the consistency of \emph{graph-based learning}, that is, learning algorithms defined with respect to geometric graphs \textcolor{red}{(TODO): references}.
@@
 Regression by spectral projection is a special case of this general setup, where one takes $\psi_k$ to be the $k$th eigenfunction of the continuum Laplacian operator $\Delta_Q$.
-In contrast to this classical approach---in which the reference measure $Q$ and resulting Laplacian $\Delta_Q$ are determined a priori---in Laplacian eigenmaps the eigenvectors of the graph Laplacian serve as the basis.
+In contrast to this classical approach---in which the reference measure $Q$ and resulting Laplacian $\Delta_Q$ is determined a priori---in Laplacian eigenmaps eigenvectors of the graph Laplacian serve as the basis.
 These eigenvectors are data-dependent objects, and adapt to the geometry of the unknown design distribution $P$ in a rich manner.
@@
 This data-dependency is appealing.
-Classical spectral projection estimators possess attractive theoretical properties for nonparametric regression---in particular, they are minimax rate-optimal for regression over H\"{o}lder and Sobolev spaces \cite{tsybakov08,johnstone2011,gine16}---but in practice suffer from some serious drawbacks.
+Classical spectral projection estimators possess attractive theoretical properties for non-parametric regression---in particular, they are minimax rate optimal for regression over H\"{o}lder and Sobolev spaces \textcolor{blue}{(Johnstone, Tsybakov, Gine)}---but in practice suffer from some serious drawbacks.
 A basic difficulty is that finding the eigenfunctions of $\Delta_Q$ is in general non-trivial, so that it may not be possible to compute the estimator in the first place.
@@
 For these reasons, such estimators are typically proposed and studied under very restrictive conditions on the design points: for instance, that they are equally spaced fixed grid points, or that they are random but uniformly distributed on the unit cube $[0,1]^d$.
-There do exist fixes to the issues just raised, for instance using nonparametric least-squares.
-We will compare Laplacian eigenmaps to nonparametric least squares in more detail later, and for now merely point out that it fundamentally changes the estimator under consideration.
+There do exist fixes to the issues just raised, for instance using non-parametric least-squares.
+We will compare Laplacian eigenmaps to non-parametric least squares in more detail later, and for now merely point out that it fundamentally changes the estimator under consideration.
 \textcolor{red}{(For Ryan + Siva): I attempted to edit this but still don't like it.}
 
-In contrast, Laplacian eigenmaps directly approximates a spectral projection method: it projects the responses onto an orthogonal basis (eigenvectors of the graph Laplacian) that approximates the smooth eigenfunctions of $\Delta_P$.
+In contrast, Laplacian eigenmaps directly approximates a spectral projection methods; it projects the responses onto a basis (eigenvectors of the graph Laplacian) that approximates the smooth eigenfunctions of $\Delta_P$.
 Laplacian eigenmaps is perfectly well-defined, and indeed straightforward to compute, when the design $P$ is unknown.
@@
 We show that when the regression function $f_0$ is smooth, in the Sobolev sense of having weak derivatives bounded in $\Leb^2(\mc{X})$ norm, Laplacian eigenmaps methods are statistically minimax optimal, for both estimation and goodness-of-fit testing.
-Our statements hold for different relations between the dimension $d$ (or $m$) and number of derivatives $s$, depending on the problem (estimation or testing), and are summarized in Tables~\ref{tbl:estimation_rates} and~\ref{tbl:testing_rates}.
+Our statements holds for different relations between the dimension $d$ (or $m$) and number of derivatives $s$, depending on the problem (estimation or testing), and are summarized in Tables~\ref{tbl:estimation_rates} and~\ref{tbl:testing_rates}.
 \begin{table}
@@
 In each case, rates hold for all $d \in \mathbb{N}$ (under Model~\ref{def:model_flat_euclidean}), and for all $m \in \mathbb{N}, 1 < m < d$ (under Model~\ref{def:model_manifold}).
-Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps estimator is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $n^{-1/2} \lesssim M \lesssim n^{1/d}$.}
+Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps estimator is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $M \gtrsim n^{-1/2}$.}
 	\label{tbl:estimation_rates}
@@
 Rates when $d > 4s$ assume that $f_0 \in L^4(P,M)$.
-Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $n^{-1/2} \lesssim M \lesssim n^{1/d}$.}
+Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $M \gtrsim n^{-1/2}$.}
 	\label{tbl:testing_rates}
@@
 \paragraph{Related Work.}
-There is an incredible amount of work on the properties of classical spectral projection methods, which go far beyond their optimality for nonparametric regression.
-We will not attempt to summarize this work, and instead refer to \citet{wasserman2006,gyorfi2006,tsybakov08,johnstone2011,gine16} and the references therein.
+There is an incredible amount of work on the properties of classical spectral projection methods, which go far beyond their optimality for non-parametric regression.
+We will not attempt to summarize these developments, and instead refer to \citet{wasserman2006,gyorfi2006,tsybakov08,johnstone2011,gine16} and the references therein.
 
 More recent work has considered regression on a fixed graph, where one treats the design points $x_1,\ldots,x_n$ as vertices in a fixed graph $G$, and carries out inference with respect to the function evaluations $(f_0(x_i))_{i = 1}^{n}$ By this point there exists a relatively mature theory describing this setting.
-Tight upper bounds have been established that certify the optimality of graph-based methods for both nonparametric estimation \citep{wang2016,hutter2016,sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}) and testing (e.g., \citet{sharpnack2010identifying,sharpnack2013b,sharpnack2013,sharpnack2015} over different ``function'' classes (in quotes because these classes really model the $n$-dimensional vector of evaluations) We call particular attention to~\citet{sadhanala16} and~\citet{sharpnack2015}, who analyze the Laplacian eigenmaps estimator and test statistic, respectively.
+Tight upper bounds have been established that certify the optimality of graph-based methods for both non-parametric estimation \citet{wang2016,hutter2016,sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}) and testing (e.g., \citet{sharpnack2010identifying,sharpnack2013b,sharpnack2013,sharpnack2015} over different ``function'' classes (in quotes because these classes really model the $n$-dimensional vector of evaluations) We call particular attention to~\citet{sadhanala16} and~\citet{sharpnack2015}, who analyze the Laplacian eigenmaps estimator and test statistic, respectively.
 This setting is quite general, because the graph need not be a geometric graph defined on a vertex set which belongs to Euclidean space.
@@
 However, as already mentioned, when the design points are random so too are the graph Laplacian eigenvectors, and grasping their properties is in general non-trivial.
-For this reason, there has not been much analysis of random design nonparametric regression using Laplacian eigenmaps.
+For this reason, there has not been much analysis of random design non-parametric regression using Laplacian eigenmaps.
 \cite{zhou2011} consider the Laplacian eigenmaps estimator, but in the semi-supervised setting, where one additionally observes unlabeled design points $X_{n + 1},\ldots,X_{n + m}$.
 Their analysis assumed that for a fixed number of labeled samples $n$, the number of unlabeled samples $m$ grows to infinity.
-In this case the eigenvectors of the graph Laplacian converge to eigenfunctions of a continuum Laplacian, and the analysis of the resulting estimator is identical to that of a classical spectral projection estimator.
-\cite{lee2016} consider the \emph{diffusion maps} estimator---which uses the eigenvectors of a different normalization of the graph Laplacian $L$---in both the supervised and semi-supervised setups.
+In this casethe eigenvectors of the graph Laplacian converge to eigenfunctions of a continuum Laplacian, and the analysis of the resulting estimator is identical to that of a classical spectral projection estimator.
+\cite{lee2016} consider the \emph{diffusion maps} estimator---which uses the eigenvectors of a different normalization of the Laplacian $L$---in both the supervised and semi-supervised setups.
 In the supervised case, they show that the diffusion maps estimator converges to the regression function as the sample size $n \to \infty$, but at a suboptimal rate.
@@
 \citet{belkin07,vonluxburg2008,singer2017, garciatrillos18} show that eigenvalue-eigenvector pairs $(\lambda,v)$ of a graph Laplacian converge to eigenvalue-eigenfunction pairs $(\lambda,\psi)$ of a limiting differential or integral operator.
-\citet{burago2014, shi2015, trillos2019, calder2019, cheng2021} build on these works, by giving finite sample bounds, rates of convergence, and making statements uniform.
+\citet{burago2014, shi2015, trillos2019, calder2019, cheng2021} build on these works, by giving finite sample bounds, rates of convergence, and making statements uniform over sets of eigenpairs.
 These results justify our intuition that Laplacian eigenmaps, which projects the responses ${\bf Y}$ onto eigenvectors of a graph Laplacian, is approximating a classical spectral projection method, which projects ${\bf Y}$ onto eigenfunctions of the limiting differential operator $\Delta_P$.
@@
 Unfortunately, the resulting rates of convergence implied by this approach are severely suboptimal, compared to the minimax rates.
-Our analysis showing that Laplacian eigenmaps achieves minimax optimal rates of convergence will use some of the results mentioned above, but will overall proceed by a very different route than the one just considered.
-
-Finally, we point out that there are other ways to use neighborhood graphs, and specifically graph Laplacians, for nonparametric regression.
+Our analysis showing that Laplacian eigenmaps achieves minimax optimal rates of converges will use some of the results mentioned above, but will proceed by a very different route than the one just given.
+
+Finally, we point out that there are other ways to use neighborhood graphs, and specifically graph Laplacians, for non-parametric regression.
 For instance, \cite{trillos2020} \textcolor{blue}{(Green 2021)} use the graph Laplacian to induce a penalty over functions $f: \{\bf X\} \to \Reals$.
@@
 \end{equation*}
-\textcolor{blue}{(Green 2021)} show that the resulting estimator is minimax optimal, but only for $s = 1$ and $d \leq 4$.
-In contrast, we show in this work that the Laplacian eigenmaps estimator is optimal for all $s$ and $d$.
+\textcolor{blue}{(Green 2021)} show that the resulting estimator is minimax optimal, but only for $s = 1$ and $d \leq 4$. In contrast, the Laplacian eigenmaps estimator is optimal for all $s$ and $d$.
 
@@
 In Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity}, we establish our main results regarding the optimality of Laplacian eigenmaps.
-Since thhe Laplacian eigenmaps estimator $\wh{f}$ is defined only at the design points $X_1,\ldots,X_n$, in Section~\ref{sec:out_of_sample} we propose a method for out-of-sample extension of $\wh{f}$, and show that it has optimal out-of-sample error.
-In Section~\ref{sec:experiments} we examine the empirical behavior of Laplacian eigenmaps, and compare it to some natural competitors for nonparametric regression.
+The Laplacian eigenmaps estimator $\wh{f}$ is defined only at the design points $X_1,\ldots,X_n$.
+In Section~\ref{sec:out_of_sample}, we propose a method for out-of-sample extension of $\wh{f}$, and show that it has optimal out-of-sample error.
+In Section~\ref{sec:experiments} we examine the empirical behavior of Laplacian eigenmaps, and compare it to some natural competitors for non-parametric regression.
 We conclude with some discussion in Section~\ref{sec:discussion}.
@@
 
-We write $\|\cdot\|$ for Euclidean, and $d_{\mc{X}}(x',x)$ for the geodesic distance between points $x$ and $x'$ on a manifold $\mc{X}$.
+Euclidean norm is $\|\cdot\|$, and $d_{\mc{X}}(x',x)$ is the geodesic distance between points $x$ and $x'$ on a manifold $\mc{X}$.
 Then for a given $\delta > 0$, $B(x,\delta)$ is the radius-$\delta$ ball with respect to Euclidean distance, whereas $B_{\mc{X}}(x,\delta)$ is the radius-$\delta$ ball with respect to geodesic distance.
-Letting $T_x(\mc{X})$ be the tangent space at a point $x \in \mc{X}$, we write $B_m(v,\delta) \subset T_x(\mc{X})$ for the radius-$\delta$ ball centered at $v \in T_x(\mc{X})$.
 
@@
 On the other hand we write $a_n = o(b_n)$ when $\lim a_n/b_n = 0$, and likewise $a_n = \omega(b_n)$ when $\lim a_n/b_n = \infty$.
-Finally $a \vee b := \max\{a,b\}$ and $a \wedge b := \min\{a,b\}$.
+Finally, we write $a \vee b := \max\{a,b\}$ and $a \wedge b := \min\{a,b\}$.
 
@@
 In this section, we begin by giving a precise definition of our framework, and the Laplacian eigenmaps methods we study.
-We then review minimax rates for nonparametric regression over Sobolev spaces.
+We then review minimax rates for non-parametric regression over Sobolev spaces.
 We pay special attention to the classical spectral projection methods which achieve these rates, since such methods are closely connected to Laplacian eigenmaps.
 
-\subsection{Nonparametric regression with random design}
+\subsection{Non-parametric regression with random design}
 \label{sec:regression_laplacian_eigenmaps}
 
-We will operate in the usual setting of nonparametric regression with random design, in which we observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$.
-The design points $X_1,\ldots,X_n$ are sampled from a distribution $P$ with support $\mc{X} \subseteq \Rd$, and the responses follow the signal plus noise model
+We will operate in the usual setting of non-parametric regression with random design, in which we observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$.
+The design points $X_1,\ldots,X_n$ are sampled from a distribution $P$ with support $\mc{X} \subseteq \Rd$, and the responses follow
 \begin{equation}
@@
 Both are derived from eigenvectors of a graph Laplacian.
-For a positive, symmetric kernel $\eta: [0,\infty) \to [0,\infty)$, and a radius parameter $\varepsilon > 0$, let $G = ([n],W)$ be the neighborhood graph formed over the design points $\{X_1,\ldots,X_n\}$, with a weighted edge $W_{ij} = \eta(\|X_i - X_j\|/\varepsilon)$ between vertices $i$ and $j$.
+For a positive, symmetric kernel $\eta: [0,\infty) \to [0,\infty)$, and a radius parameter $\varepsilon > 0$, let $G = ([n],W)$ be the neighborhood graph formed over the design points $\{X_1,\ldots,X_n\}$, with weighted edge $W_{ij} = \eta(\|X_i - X_j\|/\varepsilon)$ between points $X_i$ and $X_j$.
 Then the
@@
 
+% AG: Said elsewhere. 
+% On a different note, it is worth explicitly pointing out that $\wh{f}$ is defined only in-sample, that is, only at the design points $X_1,\ldots,X_n$.
+There is no straightforward or agreed upon way to extend $\wh{f}$ out-of-sample, and we address the problem of out-of-sample estimation in some detail later in Section~\ref{sec:out_of_sample}.
+%
+
 \subsection{Sobolev Classes}
 \label{sec:sobolev}
-We now review the definition of Sobolev classes, dividing our discussion into two cases---the flat Euclidean case and the manifold case, corresponding respectively to Models~\ref{def:model_flat_euclidean} and Model~\ref{def:model_manifold}.
+Before we give an overview of our main results, we pause to review the definition of Sobolev classes.
+We divide our discussion into two cases---the flat Euclidean case, and the manifold case, corresponding to Models~\ref{def:model_flat_euclidean} and Model~\ref{def:model_manifold}.
 
@@
 \begin{equation*}
-\int_{\mc{X}} h g = (-1)^{|\alpha|} \int_{\mc{X}} f D^{\alpha}g, \quad \textrm{for every $g \in C_c^{\infty}(\mc{X})$.}
-\end{equation*}
-If such a function $h$ exists, it is the $\alpha$th weak partial derivative of $f$, and denoted by $D^{\alpha}f := h$.
-
-\begin{definition}[Sobolev space on an open set]
-	\label{def:sobolev_space}
-	For an integer $s \geq 1$, a function $f \in L^2(\mc{X})$ belongs to the Sobolev space $H^s(\mc{X})$ if for all $|\alpha| \leq s$, the weak derivatives $D^{\alpha}f$ exists and satisfy $D^{\alpha}f \in L^2(\mc{X})$.
+\int_{\mc{X}} h g = (-1)^{|\alpha|} \int_{\mc{X}} f D^{\alpha}g, \quad \textrm{for every $g \in C_0^{\infty}(\mc{X})$.}
+\end{equation*}
+If such a function $h$ exists, it is the $\alpha$th weak partial derivative of $f$, and denoted by $D^{\alpha}f = h$.
+
+% If $f$ is $\alpha$-weakly differentiable for all $|\alpha| = j$, then $d^jf := \sum_{|\alpha| = j}D^{\alpha}f$ is the sum of all weak partial derivatives of order $j$.
+
+For an integer $s \geq 1$, the Sobolev class $H^s(\mc{X})$ consists of all those functions $f \in \Leb^2(\mc{X})$ which have weak derivatives of all orders $j = 1,\ldots,s$, all of which also belong to $\Leb^2(\mc{X})$.
 The $j$th order semi-norm for $f \in H^s(\mc{X})$ is $|f|_{H^j(\mc{X})} := \sum_{|j| = s}\|D^{\alpha}f\|_{\Leb^2(\mc{X})}$, and the corresponding norm
@@
 \end{equation*} 
-\end{definition}
 We note that $H^s(\mc{X})$ is the completion of $C^{\infty}(\mc{X})$ with respect to the $\|\cdot\|_{H^s(\mc{X})}$ norm, so that $C^{\infty}(\mc{X})$ is dense in $H^s(\mc{X})$.
@@
 To do so, we first recall the notion of partial derivatives on a manifold, which are defined with respect to a local coordinate system.
-Letting $r_1,\ldots,r_m$ be the standard basis of $\Reals^m$, for a given chart $(\phi,U)$ (meaning an open set $U \subseteq \mc{X}$, and a smooth mapping $\phi: U \to \Reals^m$) we write $\phi =: (x_1,\ldots,x_m)$ in local coordinates, meaning $x_i = r_i \circ \phi$.
+Letting $r_1,\ldots,r_m$ be the standard basis of $\Reals^m$, for a given chart $(\phi,U)$ we write $\phi = (x_1,\ldots,x_m)$ in local coordinates, meaning $x_i = r_i \circ \phi$.
 Then the partial derivative $\partial f/\partial x_i$ of a function $f$ with respect to $x_i$ at $x \in U$ is
 \begin{equation*}
-\frac{\partial f}{\partial x_i}(x) := \frac{\partial(f \circ \phi^{-1})}{\partial r_i}\bigl(\phi(x)\bigr).
+\frac{\partial f}{\partial x_i}(x) := \frac{\partial}{\partial r_i}\Big|_{\phi(x)} (f \circ \phi^{-1}).
 \end{equation*}
@@
 	A function $f \in \Leb^2(\mc{X})$ belongs to the Sobolev space $H^{s}(\mc{X})$ if for all $\abs{\alpha} \leq s$, the weak derivatives $D^{\alpha}f$ exist and satisfy  $D^{\alpha}f \in \Leb^2(\mc{X})$.
-The $j$th order semi-norm $|f|_{H^j(\mc{X})}$, the norm $\|f\|_{H^s(\mc{X})}$, and the ball $H^s(\mc{X};M)$ are all defined as in Defintion~\ref{def:sobolev_space}.
+The $j$th order Sobolev semi-norm of $f \in H^s(\mc{X})$ is $|f|_{H^j(\mc{X})} := \sum_{|j| = s}\|D^{\alpha}f\|_{\Leb^2(\mc{X})}$.
+The Sobolev norm $\|f\|_{H^s(\mc{X})}$ is defined by
+	\begin{equation}
+	\label{eqn:sobolev_space_manifold}
+	\|f\|_{H^s(\mc{X})}^2 := \|f\|_{\Leb^2(\mc{X})}^2 + \sum_{j = 1}^{s} |f|_{H^j(\mc{X})}^2.
+	\end{equation}
 \end{definition}
 The partial derivatives $D^{\alpha}f$ will depend on the choice of local coordinates, and so will the resulting Sobolev norm~$\|f\|_{H^s(\mc{X})}$.
-However, regardless of the choice of local coordinates the resulting norms will be equivalent\footnote{Recall that norms $\|\cdot\|_1$ and $\|\cdot\|_2$ on a space $\mc{F}$ are said to be equivalent if there exist constants $c$ and $C$ such that
-\begin{equation*}
-c \|f\|_1 \leq \|f\|_2 \leq C \|f\|_2 \quad \textrm{for all} f \in \mc{F}.
-\end{equation*}} 
-and so the ultimate Sobolev space $H^s(\mc{X})$ is independent of the choice of local coordinates.
+However, the ultimate Sobolev space $H^s(\mc{X})$ is independent of the choice of local coordinates.
 
@@
 \end{equation}
-As we have already alluded to, it is known~\citep{garciatrillos18} that each graph Laplacian eigenpair ($\lambda_k$,$v_k$) converges to a corresponding solution ($\lambda_k(\Delta_P)$, $\psi_k$) of~\eqref{eqn:laplacian_eigenfunction_neumann}.
+As we have already alluded to, it is known~\citep{garciatrillos18} that each graph Laplacian eigenpair ($\lambda_k$,$v_k$) converges to a corresponding solution ($\lambda_k(\Delta_P)$, $\psi_k$) to~\eqref{eqn:laplacian_eigenfunction_neumann}.
 Thus it is relevant to consider which Sobolev functions $f \in H^s(M)$ we can reconstruct using the eigenfunctions $\psi_1,\psi_2,\ldots$.
@@
 \end{equation}
-Under the conditions $p \in C^{\infty}(\mc{X})$ and $\partial\mc{X} \in C^{1,1}$, \cite{dunlop2020} show the strict inclusion $\mc{H}^{2s}(\mc{X}) \subset H^{2s}(\mc{X})$.
+Under the conditions $p \in C^{\infty}(\mc{X})$ and $\partial\mc{X} \in C^{1,1}$, \textcolor{blue}{(Thorpe 2019)} show the strict inclusion $\mc{H}^{2s}(\mc{X}) \subset H^{2s}(\mc{X})$.
 More precisely, they show that for any $s > 0$,
@@
 \label{eqn:spectral_sobolev_space_equivalence}
-\mc{H}^{2s}(\mc{X}) = \biggl\{f \in H^{2s}(\mc{X}): \frac{\partial \Delta_P^rf}{\partial {\bf n}} = 0~\textrm{on}~\partial\mc{X},~~\textrm{for all $0 \leq r \leq s - 1$} \biggr\},
-\end{equation}
-and for any $s \geq 0$, $\mc{H}^{2s + 1}(\mc{X}) = \mc{H}^{2s}(\mc{X}) \cap H^{2s + 1}(\mc{X})$.
+\mc{H}^{2s}(\mc{X}) = \biggl\{f \in H^{2s}(\mc{X}): \frac{\partial \Delta_P^rf}{\partial {\bf n}} = 0,~~\textrm{for all $0 \leq r \leq s - 1$ on $\partial\mc{X}$}\biggr\},
+\end{equation}
+where for any $s \geq 0$, $\mc{H}^{2s + 1}(\mc{X}) = \mc{H}^{2s}(\mc{X}) \cap H^{2s + 1}(\mc{X})$.
 If $P$ is uniform on $\mc{X}$, then~\eqref{eqn:spectral_sobolev_space} means that a Sobolev function $f \in H^{s}(\mc{X})$ additionally belongs to $\mc{H}^{s}(\mc{X})$ only if all its odd lower order derivatives vanish at the boundary $\partial\mc{X}$.
@@
 
-On the other hand, in the manifold model (Model~\ref{def:model_manifold}) the domain $\mc{X}$ is without boundary---precisely, every point $x \in \mc{X}$ has a neighborhood that is homeomorphic to an open set in $\Reals^m$, for instance the open ball $B(x,\delta)$ for any $\delta$ smaller than the injectivity radius $i_0$---and so boundary conditions are irrelevant.
+On the other hand, in the manifold model (Model~\ref{def:model_manifold}) the domain $\mc{X}$ is without boundary---precisely, every point $x \in \mc{X}$ has a neighborhood that is homeomorphic to an open set in $\Reals^m$---and so boundary conditions are irrelevant.
 
@@
 \label{eqn:sobolev_space_minimax_estimation_rate}
-\inf_{\wh{f}} \sup_{f_0} \Ebb\|\wh{f} - f_0\|_P^2 \asymp M^2(M^2n)^{-2s/(2s + d)};
-\end{equation}
-here infimum taken over all estimators $\wh{f}$, and the supremum over all $f_0 \in H^1(\mc{X};M)$ (first-order case) or $f_0 \in H_0^s(\mc{X};M)$ (higher-order case), and we assume $n^{-1/2} \lesssim M$.
-
-The lower bound in~\eqref{eqn:sobolev_space_minimax_estimation_rate} is due to~\citep{stone1980} (at least for the case of $M$ constant, as is most typically considered).
-The upper bound can be certified by a particular spectral projection estimator $\wt{f}$,
+\inf_{\wh{f}} \sup_{f_0} \Ebb\|\wh{f} - f_0\|_P^2 \asymp M^2(M^2n)^{-2s/(2s + d)},
+\end{equation}
+with the infimum taken over all estimators $\wh{f}$, and the supremum over all $f_0 \in H^1(\mc{X};M)$ (first-order case) or $f_0 \in H_0^s(\mc{X};M)$ (higher-order case).
+\textcolor{red}{(TODO): Need to provide a reference, for the lower bound at least.
+Is there a clean one that deals with Model~\ref{def:model_flat_euclidean}?}
+
+The upper bound in~\eqref{eqn:sobolev_space_minimax_estimation_rate} can be certified by a particular spectral projection estimator $\wt{f}$,
 \begin{equation}
@@
 The proof of Proposition~\ref{prop:spectral_series_estimation}, along with proofs of all of our results, can be found in the appendix.
-It is worth mentioning some aspects of the analysis here, because it sets the stage for the strategy we will use to analyze Laplacian eigenmaps.
+However, it is worth briefly discussing some aspects of the analysis here, because it sets the stage for the strategy we will use to analyze Laplacian eigenmaps.
 
@@
 \end{equation} 
-Typically, the null hypothesis $f_0 = f_0^{\ast} \in \mc{F}$ reflects the absence of interesting structure, and $\mc{F} \setminus  \{f_0^{\ast}\}$ is a set of smooth departures from this null.
+Typically, the null hypothesis $f_0 = f_0^{\ast} \in \mc{F}$ reflects the absence of interesting structure, and $\mc{F} \setminus  \{f_0^{\star}\}$ is a set of smooth departures from this null.
 To fix ideas, as in \citet{ingster2009} we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\ast} = 0$ and $\mc{F} = \mc{H}^s(\mc{X};M)$ is a Sobolev ball.
@@
 This analysis can be straightforwardly adapted to handle design distributions $P$ that satisfy the conditions of Model~\ref{def:model_flat_euclidean}, or to handle the case where $M$ is not $1$.
-Finally, the test \citet{ingster2009} use to certify the upper bound in~\eqref{eqn:sobolev_space_testing_critical_radius} is implicitly a spectral projection method: the test statistic is the $L^2(P_n)$ norm of a particular spectral projection estimator.\footnote{\cite{ingster2009} project the responses onto a the span of trigonometric basis functions.
+Finally, the test \citet{ingster2009} use to certify the upper bound in~\eqref{eqn:sobolev_space_testing_critical_radius} is implicitly a spectral projection method: the test statistic is the $L^2(P_n)$ norm of a particular spectral projection estimator.
+\footnote{\cite{ingster2009} project the responses onto a the span of trigonometric basis functions.
 Since $P$ is uniform on the unit cube, such functions are eigenfunctions of $\Delta_P$ when the right boundary conditions are imposed.}
@@
 When $4s \leq d$, the functions in $H^s(\Xset)$ are very irregular.
-Crucially, $H^s(\Xset)$ no longer continuously embeds into $\Leb^4(\Xset)$ when $4s \leq d$, and test statistics using the $L^2(P_n)$ norm are no longer guaranteed to have finite variance.\footnote{Note that this will not affect the analysis for estimation, because for estimation we only need to control the first two moments of $f_0$.} In fact, we have not seen any analysis which describes the minimax rate for nonparametric regression testing over Sobolev spaces in the $4s \leq d$ regime.
+Crucially, $H^s(\Xset)$ no longer continuously embeds into $\Leb^4(\Xset)$ when $4s \leq d$, and test statistics using the $L^2(P_n)$ norm are no longer guaranteed to have finite variance.\footnote{Note that this will not affect the analysis for estimation, because for estimation we only need to control the first two moments of $f_0$.} In fact, we have not seen any analysis which describes the minimax rate for non-parametric regression testing over Sobolev spaces in the $4s \leq d$ regime.
 However, if one explicitly assumes that $f_0 \in \Leb^4(\mc{X})$, then the critical radius is characterized by the dimension-free rate $\epsilon_n^2(\Leb^4(\mc{X};M)) \asymp Mn^{-1/2}$.\footnote{As a sanity check note that this is strictly worse than the rate in~\eqref{eqn:sobolev_space_testing_critical_radius}.
@@
 \paragraph{In-sample mean squared error.}
-As mentioned in our introduction, roughly speaking one of our main conclusions is that the Laplacian eigenmaps estimator $\wh{f}$ is minimax rate-optimal.
-It is worth being clear about what we do and do not mean by this statement.
+As mentioned in our main introduction, roughly speaking one of our main conclusions is that the Laplacian eigenmaps estimator $\wh{f}$ is minimax rate-optimal.
+It is worth being clear about what do and do not mean by this statement.
 We do not mean that the estimator $\wh{f}$ will match the upper bound given in~\eqref{eqn:sobolev_space_minimax_estimation_rate}, since such a statement does not make sense when the estimator is defined only at the random design points $X_1,\ldots,X_n$.
@@
 
-Additionally, we will not actually measure accuracy using the expectation of the loss.
+Secondly, we will not actually measure risk using the expectation of the loss.
 Rather, we will give a high-probability bound on $\|\cdot\|_n^2$.
@@
 Thus one cannot speak of the pointwise evaluation $f_0(X_i)$, as we do by defining our target of estimation to be $f_0(X_i)$, $i=1,\ldots,n$, until one selects \emph{representatives}.
-When $s > d/2$, every element $f$ of $H^s(\mc{X})$ admits a continuous version $f^{\ast}$, and as is standard we set this to be our favored representative.
+When $s > d/2$, every element $f$ of $H^s(\mc{X})$ admits a continuous version $f^{\ast}$, and it is standard to pick this as the favored representative.
 When $s \leq d/2$, some elements in $H^s(\mc{X})$ do not have any continuous version; however they admit a \emph{quasi-continuous} version \citep{evans15} known as the \emph{precise representative}, and we use this representative.
@@
 As previously explained, Laplacian eigenmaps is a discrete and noisy approximation to a spectral projection method using the eigenfunctions of $\Delta_P$.
-This is particularly useful when $P$ is unknown, or when the eigenfunctions of $\Delta_P$ cannot be explicitly computed.
-Our goal is to show that Laplacian eigenmaps methods are rate-optimal, notwithstanding the potential extra error incurred by this approximation.
+This is particularly useful $P$ is unknown, and the eigenfunctions of $\Delta_P$ cannot be computed.
+Our goal is to show that Laplacian eigenmaps methods are rate optimal, notwithstanding the potential extra error incurred by this approximation.
 In this section and the following one, we will see that this is indeed the case: the estimator $\wh{f}$, and a test using the statistic $\wh{T}$, achieve optimal estimation and goodness-of-fit testing rates over Sobolev classes.
@@
 \paragraph{Estimation.} When the kernel $\eta$, graph radius $\varepsilon$, and number of eigenvectors $K$ are chosen appropriately, we show in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that the estimator $\wh{f}$ achieves the minimax rate over $H^1(\mc{X};M)$, when error is measured in squared $L^2(P_n)$ norm.
-\begin{enumerate}[label=(K\arabic*)]
+\begin{enumerate}[label=(A\arabic*)]
 	\setcounter{enumi}{0}
@@
 	and we assume \smash{$\sigma_{\eta} := \frac{1}{d}\int_{\Rd} \|x\|^2 \eta(\|x\|) \,dx < \infty$}.
-\end{enumerate}
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{0}
 	\item 
@@
 	\label{eqn:eigenvector_estimation_fo} 
-	K = \min\Bigl\{\floor{(M^2n)^{d/(2 + d)}} \vee 1, n\Bigr\}.
+	K = \min\Bigl\{(M^2n)^{d/(2 + d)} \vee 1, n\Bigr\}.
 	\end{equation}
@@
 	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H^1(\mc{X},M)$.
-There are constants $c,C$ and $N$ (not depending on $f_0$, $M$ or $n$), such that the following statement holds for all $n \geq N$ and any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_fo}, then
+There are constants $c,C$ and $N$ (not depending on $f_0$), such that the following statement holds all $n \geq N$ and any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_fo}, then
 	\begin{equation}
@@
 \begin{itemize}
-	\item When $M = o(n^{-1/2})$, then computing Laplacian eigenmaps with $K = 1$ achieves the parametric rate $\|\wh{f} - f_0\|_n^2 \lesssim n^{-1}$, and the zero-estimator $\wh{f} = 0$ achieves the better rate $\|\wh{f} - f_0\|_n^2 \lesssim M^2$.
-However, we do not know what the minimax rate is in this regime.
+	\item When $M = o(n^{-1/2})$ or $M = \omega(n^{1/d})$, the rate in~\eqref{eqn:laplacian_eigenmaps_estimation_fo} looks different than~\eqref{eqn:sobolev_space_minimax_estimation_rate}.
+In particular, when $M = o(n^{-1/2})$, then computing Laplacian eigenmaps with $K = 1$ achieves the parametric rate $\|\wh{f} - f_0\|_n^2 \lesssim n^{-1}$, and the zero-estimator $\wh{f} = 0$ achieves the better rate $\|\wh{f} - f_0\|_n^2 \lesssim M^2$, but neither match the minimax rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}.
 On the other hand, when $M = \omega(n^{1/d})$, then computing Laplacian eigenmaps with $K = n$ achieves the rate $\|\wh{f} - f_0\|_n^2 \lesssim 1$, which is better than the rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}.
-This is because we are evaluating error in $L^2(P_n)$ rather than $L^2(P)$.
 However, in truth these are edge cases, which do not fall neatly into the framework of nonparametric regression.
@@
 
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{1}
+\begin{enumerate}[label=(A\arabic*)]
+	\setcounter{enumi}{2}
 	\item 
@@
 	\label{eqn:eigenvector_testing_fo}
-	K = \min\Bigl\{\floor{(M^2n)^{2d/(4 + d)}} \vee 1, n\Bigr\}.
+	K = \min\Bigl\{(M^2n)^{2d/(4 + d)} \vee 1, n\Bigr\}.
 	\end{equation}
@@
 Suppose additionally $f_0 \in H^1(\mc{X},M)$, and that $d < 4$.
-Then there exist constants $C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_fo}, and if $f_0$ satisfies
+Then there exist constants $C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_fo}, and if $f_0$ satisfies
 	\begin{equation}
@@
 	\item As mentioned previously, when $d \geq 4$ the first order Sobolev space $H^1(\mc{X})$ does not continuously embed into $\Leb^4(\mc{X})$, and we do not know the optimal rates for regression testing over $H^1(\mc{X},M)$.
-On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then the Laplacian eigenmaps test with $K = n$, has small type II error whenever $\|f_0\|_P^2 \gtrsim Mn^{-1/2}$.
-Note that when $K = n$ the Laplacian eigenmaps test statistic is nothing but the squared $L^2(P_n)$ norm of ${\bf Y}$, $\wh{T} = \|{\bf Y}\|_n^2$.
-See \textcolor{blue}{(Green et al.
-2021)} for details.
-	\item As in the estimation setting, the range of Sobolev ball radii $n^{-1/2} \lesssim M \lesssim n^{1/d}$ for which Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} implies that $\varphi$ is a rate-optimal test covers all those cases for which the critical radius $\epsilon(H^1(\mc{X};M))$ is both $\Omega(1/n)$ and $O(1)$.
+On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then the Laplacian eigenmaps test, with $K = n$, has small type II error whenever $\|f_0\|_P^2 \gtrsim Mn^{-1/2}$.
+\textcolor{red}{(TODO): Return to make sure the dependence on $M$ is correct.}
+	\item Similar to the estimation setting, the range of Sobolev ball radii $n^{-1/2} \lesssim M \lesssim n^{1/d}$ covers all those cases for which the critical radius $\epsilon(H^1(\mc{X};M))$ is both $\Omega(1/n)$ and $O(1)$.
 \end{itemize}
@@
 In order to show that $\wh{f}$ is an optimal estimator over $H_0^s(\mc{X};M)$, we will require that $\varepsilon$ be meaningfully larger than the lower bound in~\ref{asmp:parameters_estimation_fo}.
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{2}
+\begin{enumerate}[label=(A\arabic*)]
+	\setcounter{enumi}{3}
 	\item 
 	\label{asmp:parameters_estimation_ho}
-	For constants $c_0$ and $C_0$, the graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy
+	The graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy
 	\begin{equation}
 	\label{eqn:radius_ho}
-	C_0\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/d}, (M^2n)^{-1/(2(s - 1) + d)}\biggr\} \leq \varepsilon \leq c_0\min\{1, K^{-1/d}\}
+	\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/d}, (M^2n)^{-1/(2(s - 1) + d)}\biggr\} \ll \varepsilon \ll \min\{1, K^{-1/d}\}
 	\end{equation}
@@
 	\begin{equation*}
-	K = \min\Bigl\{\floor{(M^2n)^{d/(2s + d)}} \vee 1,n\Bigr\}
+	K \asymp (M^2n)^{2d/(2s + d)}.
 	\end{equation*}
 \end{enumerate}
-Crucially, when $n$ is sufficiently large the two conditions in~\ref{asmp:parameters_estimation_ho} are guaranteed to not be mutually exclusive.
-This is because so long as $M^2 = \omega(n^{-1})$ then $(M^2n)^{-2/(2(s - 1) + d)} = o((M^2n)^{-2/(2s + d)})$, regardless of $s$ and $d$.
+Crucially, the two conditions in~\ref{asmp:parameters_estimation_ho} are not mutually exclusive.
+So long as $M^2 = \omega(n^{-1})$ then $(M^2n)^{-2/(2(s - 1) + d)} = o((M^2n)^{-2/(2s + d)})$, regardless of $s$ and $d$.
 \begin{theorem}
@@
 	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H_0^s(\mc{X},M)$ and $p \in C^{s - 1}(\mc{X})$.
-There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_ho}, then
+There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_ho}, then
 	\begin{equation}
@@
 \end{theorem}
-Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that in the flat Euclidean setting Laplacian eigenmaps is an in-sample minimax rate-optimal estimator over Sobolev classes, for all values of $s$ and $d$.
+Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that in the flat Euclidean setting Laplacian eigenmaps is an in-sample minimax rate optimal estimator over Sobolev classes, for all values of $s$ and $d$.
 Some other remarks:
@@
 \paragraph{Testing.} The test $\varphi$ can adapt to the higher-order smoothness of $f_0$, when $\varepsilon$ and $K$ are chosen correctly.
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{3}
+\begin{enumerate}[label=(A\arabic*)]
+	\setcounter{enumi}{4}
 	\item 
@@
 	\label{eqn:eigenvector_testing_ho}
-	K = \min\Bigl\{\floor{(M^2n)^{2d/(4 + d)}} \vee 1, n\Bigr\}.
+	K \asymp \min\Bigl\{(M^2n)^{2d/(4s + d)} \vee 1, n\Bigr\}.
 	\end{equation}
 \end{enumerate}
-When $d \leq 4$, for any value of $s \in \mathbb{N}$ when $n$ is sufficiently large it is possible to choose $\varepsilon$ and $K$ such that both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho} are satisfied, and our next theorem establishes that in this situation $\varphi$ is an optimal test.
+When $d \leq 4$, for any value of $s \in \mathbb{N}$ it is possible to choose $\varepsilon$ and $K$ such that both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho} are satisfied, and our next theorem establishes that in this situation $\varphi$ is an optimal test.
 \begin{theorem}
@@
 Suppose additionally $f_0 \in H_0^s(\mc{X},M)$, that $p \in C^{s-1}(\mc{X})$, and that $d \leq 4$.
-Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n \geq N$: if the Laplacian eigenmaps test $\varphi$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_ho}, and if $f_0$ satisfies
+Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_ho}, and if $f_0$ satisfies
 	\begin{equation}
@@
 Similarly to the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that $\varphi$ is a minimax optimal test over $H_0^s(\mc{X})$ when $n^{-1/2} \lesssim M^2 \lesssim n^{1/d}$.
+
 However, unlike the first-order case, when $4 < d < 4s$ the minimax testing rate over $H_0^s(\mc{X})$ is still on the order of $M^2(M^2n)^{-4s/(4s + d)}$.
@@
 	Under the same setup as Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, but with $4 < d < 4s$.
-If the Laplacian eigenmaps test $\varphi$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, number of eigenvectors $K$ satisfying~\eqref{eqn:eigenvector_testing_ho}, and $\varepsilon = (M^2n)^{-1/(2(s - 1) + d)}$, and if
+If the Laplacian eigenmaps test $\varphi$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel}, number of eigenvectors $K$ satisfying~\eqref{eqn:eigenvector_testing_ho}, and $\varepsilon = (M^2n)^{-1/(2(s - 1) + d)}$, and if
 	\begin{equation}
@@
 \end{theorem}
-Note that as a consequence of Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, if we choose $K = n^{-2s/(2s + d)}$ then $\varphi$ must have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2s + d)}$.
+Note that by Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, choosing $K = n^{-2s/(2s + d)}$ as is optimal for estimation, $\varphi$ must have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2s + d)}$.
 Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} shows that $\varphi$ can achieve better, but still not optimal, rates.
-As a technical matter, the problem is that when $d > 4$ there do not exist any choices of $\varepsilon$ and $K$ which satisfy both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho}, and as a result we cannot optimally balance (our upper bound on) testing bias and variance (defined momentarily in~\eqref{eqn:testing_biasvariance}).
+As a technical matter, the problem is that there do not exist any choices of $\varepsilon$ and $K$ which satisfy both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho}, and as a result we cannot optimally balance (our upper bound on) testing bias and variance (defined momentarily in~\eqref{eqn:testing_biasvariance}).
 Although we suspect $\varphi$ is truly suboptimal when $d > 4$, technically speaking~\eqref{eqn:testing_biasvariance} gives only an upper bound on testing bias, and thus we cannot rule out that the test $\varphi$ is optimal for all $4 < d < 4s$.
@@
 Of course when the index $K$ increases with $n$, as is necessary to optimally balance bias and variance, the issue only gets worse.
-Clearly, as a method for regression, the rate of convergence of Laplacian eigenmaps is much better than the rate implied by (what is currently known about) the concentration of individual eigenvectors around their continuum limits.
+Clearly, as a method for regression, the rate of convergence of Laplacian eigenmaps is much better than that implied by (what is currently known about) the concentration of individual eigenvectors around their continuum limits.
 
@@
 We now outline the high-level strategy we follow when proving each of Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}.
-We analyze the estimation error of $\wh{f}$, and the testing error of $\wh{\varphi}$, by first conditioning on the design points $X_1,\ldots,X_n$ and deriving \emph{design-dependent} bias and variance terms.
+We analyze the estimation error of $\wh{f}$, and the testing error of $\wh{\varphi}$, by first conditioning on the design points ${\bf X}$ and deriving \emph{design-dependent} bias and variance terms.
 For estimation, we have that with probability at least $1 - \exp(-K)$,
@@
 	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f \in H_0^s(\mc{X})$ and $p \in C^{s - 1}(\mc{X})$.
-Then there exist constants $c$ and $C$ that do not depend on $f_0$ or $n$ such that the following statement holds for any $\delta \in (0,1)$: if $\eta$ satisfies~\ref{asmp:kernel_flat_euclidean} and $Cn^{-1/(2(s - 1) + d)} < \varepsilon < c$, then
+Then there exist constants $c$ and $C$ that do not depend on $f_0$ or $n$ such that the following statement holds for any $\delta \in (0,1)$: if $\eta$ satisfies~\ref{asmp:kernel} and $Cn^{-1/(2(s - 1) + d)} < \varepsilon < c$, then
 	\begin{equation}
@@
 We now summarize the techniques used to prove Proposition~\ref{prop:graph_seminorm_ho}, which will help explain what role the conditions on $f_0$,$p$ and $\varepsilon$ play.
-To upper bound $\dotp{L_{n,\varepsilon}^sf}{f}_n$ in terms of $\|f\|_{H^s(\mc{X})}^2$, we introduce an intermediate quantity: the \emph{order-s non-local Sobolev seminorm} $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$.
+To upper bound $\dotp{L_{n,\varepsilon}^sf,f}_n$ in terms of $\|f\|_{H^s(\mc{X})}^2$, we introduce an intermediate quantity: the \emph{order-s non-local Sobolev seminorm} $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$.
 This seminorm is defined with respect to $L_{P,\varepsilon}$, which is a non-local approximation to $\Delta_P$,
@@
 Specifically, $\dotp{L_{n,\varepsilon}^s f}{f}_n$ is a $V$-statistic, meaning it is the sum of an unbiased estimator of $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$ (in other words, a $U$-statistic) plus some higher-order, pure bias terms.
-We show that these pure bias terms are negligible when $\varepsilon = \omega(n^{-1/(2(s - 1) + d)})$.
+We show that these pure bias terms are negligible when $\varepsilon = \omega(n^{-1/(2r + d)})$.
 	\item For $x$ sufficiently in the interior of $\mc{X}$, we show that $L_{P,\varepsilon}^jf(x) \to \sigma_{\eta}^j \Delta_P^jf(x)$ as $\varepsilon \to 0$.
@@
 	\item Finally, we combine the results of previous two steps to deduce an upper bound on $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$ in terms of the squared Sobolev norm $\|f\|_{H^s(\mc{X})}^2$.
-Roughly speaking, when $s$ is odd, $\dotp{L_{P,\varepsilon}^sf}{f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}^jf) \approx \sigma_{\eta}^{2j}E_{P,\varepsilon}(\Delta_P^jf)$, whereas when $s$ is even $\dotp{L_{P,\varepsilon}^sf}{f}_P = \|L_{P,\varepsilon}L_{P,\varepsilon}^{j}f\|_{P}^2 \approx \sigma_{\eta}^{2j}\|L_{P,\varepsilon} \Delta_Pf\|_P^2$.
-Reasoning in this way, we can translate estimates of $L_{P,\varepsilon}^jf$ into an upper bound on the order-$s$ non-local Sobolev seminorm, even though $s > j$.
+Note that when $s$ is odd,  $\dotp{L_{P,\varepsilon}^sf}{f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}^jf)$.
+This explains how we can translate estimates of $L_{P,\varepsilon}^jf(x)$ into an upper bound on the order-$s$ non-local Sobolev seminorm, even though $s > j$.
 \end{itemize}
@@
 It is worth pointing out that we do not try to show $L_{P,\varepsilon}^sf(x) \to \Delta_{P}^sf(x)$.
-This may seem like a natural first step towards a simple proof that $\dotp{L_{P,\varepsilon}^sf}{f}_{P} \to \dotp{\Delta_P^sf}{f}_{P}$.
+This may seem like a natural first step towards establishing $\dotp{L_{P,\varepsilon}^sf}{f}_{P} \to \dotp{\Delta_P^sf}{f}_{P}$.
 The problem is that $\Delta_P^s$ is an order-$2s$ differential operator, whereas we assume that $f$ has only $s$ bounded derivatives.
@@
 	Suppose Model~\ref{def:model_flat_euclidean}.
-Then there exist constants $c$ and $C$ such that the following statement holds: if $\eta$ satisfies~\ref{asmp:kernel_flat_euclidean} and $C(\log n/n)^{1/d} < \varepsilon < c$, then
+Then there exist constants $c$ and $C$ such that the following statement holds: if $\eta$ satisfies~\ref{asmp:kernel} and $C(\log n/n)^{1/d} < \varepsilon < c$, then
 	\begin{equation}
@@
 On the other hand, when $s > 1$, we can show Laplacian eigenmaps is optimal only when $\varepsilon = \omega(n^{-c})$ for some $c < 1/d$.
-For such a choice of $\varepsilon$, the average degree in $G$ will grow polynomially in $n$ as $n \to \infty$, and computing eigenvectors of the Laplacian of a graph will be more computationally intensive than if the graph were sparse \textcolor{red}{(reference)}.
+For such a choice of $\varepsilon$, the average degree in $G$ will grow polynomially in $n$ as $n \to \infty$.
+Computing eigenvectors of the Laplacian of a graph becomes more computationally intensive as the average degree grows \textcolor{red}{(reference)}.
 Thus Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and~\ref{thm:laplacian_eigenmaps_estimation_ho} can be seen as revealing a tradeoff between statistical and computational efficiency; although to be clear, we have no theoretical evidence that Laplacian eigenmaps \emph{fails} to adapt to higher-order smoothness when $\varepsilon \asymp (\log n/n)^{1/d}$---we simply cannot prove that it succeeds.
@@
 By now there exist various methods see (e.g., the seminal papers of \citet{spielman2011,spielman2013,spielman2014}, or the overview by \citet{vishnoi2012} and references therein) to efficiently remove many edges from the graph $G$ while only slightly perturbing the spectrum of the Laplacian.
-Specifically such algorithms take as input a parameter $\sigma \geq 1$, and return a sparser graph $\wt{G}$, $E(\wt{G}) \subseteq E(G)$, with a Laplacian $\wt{L}_{n,\varepsilon}$ satisfying
-\begin{equation*}
-\frac{1}{\sigma} \cdot u^{\top} \wt{L}_{n,\varepsilon} u \leq u^{\top} L_{n,\varepsilon} u \leq \sigma \cdot u^{\top} \wt{L}_{n,\varepsilon}u \quad \textrm{for all $u \in \Reals^n$.}
-\end{equation*}
-Let $\wt{f}$ be the Laplacian eigenmaps estimator computed using the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ .
-Because $\wt{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wt{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$, and consequently much faster to compute $\wt{f}$ than $\wh{f}$ \textcolor{red}{(reference needed)}.
-Statistically speaking, letting $\wt{\lambda}_k$ be the $k$th eigenvalue of $\wc{L}_{n,\varepsilon}$, we have that conditional on $X_1,\ldots,X_n$,
-\begin{equation*}
-\|\wt{f} - f_0\|_n^2 \leq \frac{\dotp{\wt{L}_{n,\varepsilon}^s f_0}{f_0}_n}{\wt{\lambda}_{K + 1}^s} + \frac{5K}{n} \leq \sigma^{2s} \frac{\dotp{\wt{L}_{n,\varepsilon}^s f_0}{f_0}_n}{\wt{\lambda}_{K + 1}^s} + \frac{5K}{n},
+Specifically such algorithms take as input a parameter $\sigma \geq 1$, and return a sparser graph $\wc{G}$, $E(\wc{G}) \subseteq E(G)$, with a Laplacian $\wc{L}_{n,\varepsilon}$ satisfying
+\begin{equation*}
+\frac{1}{\sigma} \cdot u^{\top} \wc{L}_{n,\varepsilon} u \leq u^{\top} L_{n,\varepsilon} u \leq \sigma \cdot u^{\top} \wc{L}_{n,\varepsilon}u \quad \textrm{for all $u \in \Reals^n$.}
+\end{equation*}
+Let $\wc{f}$ be the Laplacian eigenmaps estimator computed using the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ .
+Because $\wc{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wc{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$, and consequently much faster to compute $\wc{f}$ than $\wh{f}$ \textcolor{red}{(reference needed)}.
+Statistically speaking, letting $\wc{\lambda}_k$ be the $k$th eigenvalue of $\wc{L}_{n,\varepsilon}$, we have that conditional on $X_1,\ldots,X_n$,
+\begin{equation*}
+\|\wc{f} - f_0\|_n^2 \leq \frac{\dotp{\wc{L}_{n,\varepsilon}^s f_0}{f_0}_n}{\wc{\lambda}_{K + 1}^s} + \frac{5K}{n} \leq \sigma^{2s} \frac{\dotp{\wc{L}_{n,\varepsilon}^s f_0}{f_0}_n}{\wc{\lambda}_{K + 1}^s} + \frac{5K}{n},
 \end{equation*}
 with probability at least $1 - \exp(-K)$.
-Consequently $\wt{f}$ has $L^2(P_n)$-error of at most $\sigma^{2s}$ times our upper bound on the error of $\wh{f}$, and for any choice of $\sigma$ that is constant in $n$ the estimator $\wt{f}$ will also be rate-optimal.
+Consequently $\wc{f}$ has $L^2(P_n)$-error of at most $\sigma^{2s}$ times our upper bound on the error of $\wh{f}$, and for any choice of $\sigma$ that is constant in $n$ the estimator $\wc{f}$ will also be rate-optimal.
 
@@
 For one thing, they are designed to work when $\sigma$ is much larger than $1$, whereas in order for $\wc{f}$ to be rate-optimal setting $\sigma$ to be any constant greater than $1$, say $\sigma = 2$, is sufficient.
-Additionally, edge sparsification algorithms are traditionally designed to work in the worst-case, where no assumptions are made on the structure of the graph $G$.
-But the geometric graphs we consider in this paper exhibit a special structure, in which very roughly speaking no single edge is a bottleneck.
+Additionally, edge sparsification algorithms are traditionally designed to work in the worst-case, that is to work regardless of the structure of the graph.
+But geometric graphs such as $G$ exhibit a special structure, in which very roughly speaking, no single edge is a bottleneck.
 As pointed out by~\citet{sadhanala16b}, in this special case there exist far simpler and faster methods for sparsification, which at least empirically seem to do the job.
@@
 In this section we consider the manifold setting, where $(X_1,Y_1),\ldots,(X_n,Y_n)$ are observed according to Model~\ref{def:model_manifold}.
-A theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+A theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
 We build on this work by showing that when $f_0 \in H^s(\mc{X})$ and $P$ is supported on a manifold, Laplacian eigenmaps achieve the sharper minimax estimation and testing rates reviewed in Section~\ref{subsec:minimax_rates_sobolev}.
@@
 \subsection{Laplacian eigenmaps error rates under the manifold hypothesis}
-Unlike in the flat-Euclidean case, since Model~\ref{def:model_manifold} assumes that $\mc{X}$ is boundaryless it is easy to deal with the first-order $(s = 1)$ and higher-order $(s > 1)$ cases all at once.
+Unlike in the flat-Euclidean case, because in Model~\ref{def:model_manifold} we assume that $\mc{X}$ is boundaryless it is easy to deal with the first-order $(s = 1)$ and higher-order $(s > 1)$ cases all at once.
 A more important distinction between the results of this section and those of Section~\ref{sec:minimax_optimal_laplacian_eigenmaps} is that we will establish Laplacian eigenmaps is optimal only when the regression function $f_0 \in H^s(\mc{X};M)$ for $s \leq 3$.
@@
 
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{4}
+\begin{enumerate}[label=(A\arabic*)]
+	\setcounter{enumi}{5}
 	\item 
@@
 	\label{asmp:parameters_estimation_manifold}
-	For a constant $C_0$, the graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy
+	The graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy
 	\begin{equation}
 	\label{eqn:radius_estimation_manifold}
-	C_0\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/m}, n^{-1/(2(s - 1) + m)}\biggr\} \leq \varepsilon \leq \min\{i_0, K^{-1/m}\}
+	\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/m}, n^{-1/(2(s - 1) + m)}\biggr\} \ll \varepsilon \ll \min\{1, K^{-1/m}\}
 	\end{equation}
@@
 	\begin{equation*}
-	K = \min\Bigl\{\floor{(M^2n)^{m/(2s + m)}} \wedge 1,n \Bigr\}.
+	K \asymp (M^2n)^{2m/(2s + m)}.
 	\end{equation*}
@@
 Likewise, to construct a minimax optimal test using $\wh{T}$, we choose $\varepsilon$ and $K$ as in~\ref{asmp:parameters_testing_fo}, except with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{5}
+\begin{enumerate}[label=(A\arabic*)]
+	\setcounter{enumi}{7}
 	\item 
@@
 	\begin{equation*}
-	K = \min\Bigl\{\floor{(M^2n)^{2m/(4s + m)}} \wedge 1,n \Bigr\}.
+	K \asymp C_2 (M^2n)^{2m/(4s + m)}
 	\end{equation*}
@@
 	\item Unlike in the full-dimensional case, our upper bounds on the estimation and testing error of Laplacian eigenmaps match the minimax rate only when $s \leq 3$.
-Our upper bounds when $s \geq 4$ follow from the embedding $H^s(\mc{X};M) \subset H^{3}(\mc{X};M)$, i.e they match the rates we get just by assuming that the $3$rd order derivative is bounded, and are clearly suboptimal.
+Our upper bounds when $s \geq 4$ follow from the embedding $H^s(\mc{X};M) \subset H^{3}(P;M)$, i.e they match the rates we get just by assuming that the $3$rd order derivative is bounded, and are clearly suboptimal.
 
@@
 	Suppose Model~\ref{def:model_manifold}.
-Then there exist constants $c$ and $C$ such that the following statement holds: if $\eta$ satisfies~\ref{asmp:kernel_manifold} and $C(\log n/n)^{1/m} < \varepsilon < c$, then
+Then there exist constants $c$ and $C$ such that the following statement holds: if $\eta$ satisfies~\ref{asmp:kernel} and $C(\log n/n)^{1/d} < \varepsilon < c$, then
 	\begin{equation}
@@
 We show that the out-of-sample error $\|T_{n,h}\wc{f} - f_0\|_P^2$ can be upper bounded by three terms--- (a constant times) the in-sample error $\|\wc{f} - f_0\|_n^2$, and variance and bias terms that arise naturally in the analysis of kernel smoothing over noiseless data..
-We shall assume the following conditions on $\psi$ and $h$.
+We shall assume the following conditions on $\psi$.
 \begin{enumerate}[label=(K\arabic*)]
-	\setcounter{enumi}{2}
+	\setcounter{enumi}{1}
 	\item
@@
 	\begin{equation*}
-	\int_{-\infty}^{\infty} \psi(|z|) \,dz = 1.
-	\end{equation*}
-\end{enumerate}
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{6}
-	\item
-	\label{asmp:bandwidth}
-	For constants $c_0$ and $C_0$, the bandwidth parameter $h$ satisfies
-	\begin{equation*}
-	C_0\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq h \leq c_0.
+	\int_{-\infty}^{\infty} \psi(|s|) \,ds = 1.
+	\end{equation*}
+	Finally, if $s > 1$ then $\psi$ is an order-$s$ kernel, meaning that it satisfies
+	\begin{equation*}
+	\int_{-\infty}^{\infty} \psi(|s|) = 1,~~\int_{-\infty}^{\infty} t^j \psi(|t|) \,dt = 0~~\textrm{for $j = 1,\ldots,s + d - 2$},~~\textrm{and}~~\int_{-\infty}^{\infty} t^{s + d - 1} \psi(|t|) \,dt < \infty.
 	\end{equation*}
@@
 	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $\wc{f} \in L^2(P_n)$, $f_0 \in H^1(\mc{X})$ and $p \in C^1(\mc{X})$.
-If the kernel smoothing estimator $T_{h,n}\wc{f}$ is computed with kernel $\psi$ satisfying~\ref{asmp:kernel} and bandwidth $h$ satisfying~\ref{asmp:bandwidth}, it holds that
+If the kernel smoothing estimator $T_{h,n}\wc{f}$ is computed with kernel $\psi$ satisfying~\ref{asmp:kernel}, it holds that
 	\begin{equation}
@@
 Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data, by a factor of $h^2$.
-On the other hand the bias term is typical.
-When $\psi$ is an order-$s$ kernel, a standard analysis shows that the $\|T_{h,P}f_0 - f_0\|_P^2 \lesssim \varepsilon^{2s}$.
-\begin{enumerate}[label=(K\arabic*)]
-	\setcounter{enumi}{3}
-	\item
-	\label{asmp:ho_kernel}
-	The kernel function $\psi$ is an order-$s$ kernel, meaning that it satisfies
-	\begin{equation*}
-	\int_{-\infty}^{\infty} \psi(|z|) \,dz = 1, \quad \int_{-\infty}^{\infty} z^j \psi(|z|) \,dz = 0 ~~\textrm{for}~j = 1,\ldots, s + d - 2, \quad \textrm{and}~ \int_{-\infty}^{\infty} z^{s + d - 1} \psi(|z|) \,dz < \infty.
-	\end{equation*}
-\end{enumerate}
+On the other hand the bias term is typical, and a standard analysis shows that the $\|T_{h,P}f_0 - f_0\|_P^2 \lesssim \varepsilon^{2s}$ when $\psi$ is an order-$s$ kernel.
 
@@
 \end{equation}
-\eqref{eqn:kernel_smoothing_insample2} tells us that the additional error incurred by passing a kernel smoother over an in-sample estimator $\wc{f}$ is negligible compared to the minimax rate of estimation.
-Consequently, if $\wc{f}$ converges at the minimax rate in $L^2(P_n)$, then $T_{n,h}\wc{f}$ will converge at the minimax rate in $L^2(P)$.
-It follows immediately from Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} (when $s = 1$) or Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} (when $s > 1$), that $T_{h,n}\wh{f}$ achieves the optimal rate of convergence in $L^2(P)$.
+This demonstrates that the additional error incurred by passing a kernel smoother over an in-sample estimator $\wc{f}$ is negligible compared to the minimax rate of estimation.
+So if $\wc{f}$ converges at the minimax rate in $L^2(P_n)$, then $T_{n,h}\wc{f}$ will converge at the minimax rate in $L^2(P)$.
+It follows from Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} (when $s = 1$) or Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} (when $s > 1$), that $T_{h,n}\wh{f}$ achieves the optimal rate of convergence in $L^2(P)$.
 
@@
 	\label{thm:laplacian_eigenmaps_estimation_out_of_sample}
-	Suppose Model~\ref{def:model_flat_euclidean}.
-There exist constants $c$, $C$, and $N$ that do not depend on $f_0$ or $n$ such that each the following statements hold with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\} - Ch^d\exp\{-cnh^d\}$,  for all $n \geq N$ and for any $\delta \in (0,1)$.
-	\begin{itemize}
-		\item If $f_0 \in H^1(\mc{X};M)$, the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_fo}, and the out-of-sample extension $T_{h,n}\wh{f}$ is computed with bandwidth $h = n^{-1/d}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel}, then
-		\begin{equation*}
-		\|T_{h,n}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2s/(2s + d)}.
-		\end{equation*}
-		\item If $f_0 \in H_0^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$ for some $s \in \mathbb{N}, s > 1$, and the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_ho}, and the out-of-sample extension $T_{h,n}\wh{f}$ is computed with bandwidth $h = n^{-1/(2(s - 1) + d)}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel} and~\ref{asmp:ho_kernel}, then
-		\begin{equation*}
-		\|T_{h,n}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2s/(2s + d)}.
-		\end{equation*}
-	\end{itemize}
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that for some $s \in \mathbb{N}$, the regression function $f_0 \in H^s(\mc{X};M)$ (if $s = 1$) or $f_0 \in H_0^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$ (if $s > 1$).
+There exist constants $c$, $C$, and $N$ that do not depend on $f_0$, $n$, or $M$, such that the following statement holds for all $n \geq N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_ho}, and the out-of-sample extension $T_{h,n}\wh{f}$ is computed with bandwidth $h = n^{-1/(2(s - 1) + d)}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel} and
+	\begin{equation*}
+	\|T_{h,n}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2s/(2s + d)},
+	\end{equation*}
+	with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\} - Ch^d\exp\{-cnh^d\}$. 
 \end{theorem}
+
 Some remarks:
 \begin{itemize}
-	\item Since $T_{n,h}\wh{f}$ is defined out-of-sample, we can use sample splitting or cross validation methods to tune hyperparameters, which we could not do for the original estimator $\wh{f}$.
+	\item Since $T_{n,h}\wh{f}$ is defined out-of-sample, we can use sample splitting or cross validation methods to tune hyperparameters.
 For instance, we can (i) split the sample into two halves, (ii) use the first half to compute $T_{h,n}\wh{f}$ for various values of $\varepsilon$, $h$, and $K$, (iii) choose the optimal values of these three hyperparameters by minimizing error on the held out set.
-Practically speaking, cross-validation is one of the most common approaches to choosing hyperparameters.
+Practically speaking, cross-validation is one of the most common approaches to choosing hyperparameters, and is  .
 Theoretically, it is known \textcolor{red}{(references)} that choosing hyper-parameters through sample splitting can result in estimators that optimally adapt to the order of regularity $s$.
-In other words, it leads to estimators that are rate-optimal (up to $\log n$ factors), even when $s$ is unknown.
+In other words, it leads to estimators that are rate-optimal (up to $\log(n)$ factors), even when $s$ is unknown.
 Similar arguments should imply that $T_{h,n}\wh{f}$ is adaptive in this sense when $\varepsilon,h$ and $K$ are chosen by sample splitting.
 \textcolor{red}{(TODO): Siva would like me to fill in some details?}
-	\item There exist many other approaches to extending a function $f$ using only evaluations $\{f(X_1),\ldots,f(X_n)\}$.
-We consider extension by kernel smoothing because it is a simple and statistically optimal procedure that does not require any knowledge of the domain $\mc{X}$ or distribution $P$---as we have argued, this latter property is one of the main selling points of Laplacian eigenmaps as a tool for nonparametric regression.
-
-	We now comment on a few alternative methods. One such approach is minimum norm interpolation. Here one defines a normed space $(\mc{F},\|\cdot\|_{\mc{F}})$ and then solves the optimization problem
+	\item We consider extension by kernel smoothing because it is a simple and statistically optimal procedure that does not require any knowledge of the domain $\mc{X}$ or distribution $P$---as we have argued, this latter property is one of the main selling points of Laplacian eigenmaps as a tool for nonparametric regression.
+
+	That being said, there exist many other approaches to extending (or interpolating) a function $f$ using only evaluations $\{f(X_1),\ldots,f(X_n)\}$, and we now comment on a few of these alternatives.
+One approach is to perform minimum norm interpolation, where one defines a normed space $(\mc{F},\|\cdot\|_{\mc{F}})$ and then solves the optimization problem
 	\begin{equation*}
@@
 This is in one sense a strength, since they can be used to extrapolate any estimator defined only in-sample.
-But it is also potentially a weakness.
-Each of these methods have their own approximation and estimation errors (bias and variance) which can be fundamentally different than those of Laplacian eigenmaps, and there is a danger that in extrapolating the Laplacian eigenmaps estimator in this way we are taking the ``worst of both worlds''.
+But it also potentially a weakness.
+Each of these methods have their own statistical and approximation errors (bias and variance) which can be fundamentally different than those of Laplacian eigenmaps, and there is a danger that in extrapolating the Laplacian eigenmaps estimator in this way we are taking the ``worst of both worlds''.
 Our theory shows that when the data model is Model~\ref{def:model_flat_euclidean} and we perform extrapolation by kernel smoothing, this is not a problem, at least in a minimax sense.
@@
 For example, suppose instead of knowing $\psi_1,\psi_2,\ldots$, we had access only to eigenfunctions $\phi_1,\phi_2,\ldots$ of an unweighted Laplace-Beltrami operator $\Delta$.
-Then letting $\Phi_K = \mathrm{span}\{\phi_1,\ldots,\phi_K\}$, the least-squares estimator and test statistic
+Then letting $\Phi_K = \mathrm{span}\{f_1,\ldots,f_K\}$, the least-squares estimator and test statistic
 \begin{equation*}
@@
 using the estimator $T_{h,n}Y$ (possibly for a different choice of $h$).
-In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{h,n}\wh{f}\|_P^2$ is of a strictly lower order than $\inf_{h'} \Ebb\|f_0 - T_{h',n}Y\|_P^2$.
+In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{h,n}\wh{f}\|_P^2$ is stochastically dominated by $\inf_{h'} \Ebb\|f_0 - T_{h',n}Y\|_P^2$.
 This is possible because Laplacian eigenmaps induces a completely different bias than kernel smoothing.

### 2021-04-28 manuscripts/laplacian_eigenmaps/appendix.tex

@@
 In this section, we adopt the fixed design perspective; or equivalently, condition on $X_i = x_i$ for $i = 1,\ldots,n$.
-Let $G = \bigl([n],W\bigr)$ be a fixed graph on $\{1,\ldots,n\}$ with Laplacian matrix $L = \frac{1}{n}\sum_{k = 1}^{n}\lambda_k v_k v_k^{\top}$; the eigenvectors have unit empirical norm, $\|v_k\|_n^2 = 1$.
+Let $G = \bigl([n],W\bigr)$ be a fixed graph on $\{1,\ldots,n\}$ with Laplacian matrix $L = \sum_{k = 1}^{n}\lambda_k v_k v_k^{\top}$; the eigenvectors have unit empirical norm, $\|v_k\|_n^2 = 1$.
 The randomness thus all comes from the responses
@@
 \label{eqn:fixed_graph_regression_model}
-Y_i = f_{0}(x_i) + \varepsilon_i
-\end{equation}
-where the noise variables $\varepsilon_i$ are independent $N(0,1)$.
+Y_i = f_{0}(x_i) + w_i
+\end{equation}
+where the noise variables $w_i$ are independent $N(0,1)$.
 In the rest of this section, we will mildly abuse notation and write $f_0 = (f_0(x_1),\ldots,f_0(x_n)) \in \Reals^n$.
@@
 	\label{lem:fixed_graph_estimation}
-	For any $s \in \mathbb{N} \setminus \{0\}$, the Laplacian eigenmaps estimator $\wh{f}$ of~\eqref{eqn:laplacian_eigenmaps_estimator} satisfies
+	For any integer $s > 0$, and any integer $0 \leq K \leq n$, the Laplacian eigenmaps estimator $\wh{f}$ of~\eqref{eqn:laplacian_eigenmaps_estimator} satisfies
 	\begin{equation}
 	\label{eqn:fixed_graph_estimation}
-	\|\wh{f} - f_0\|_n^2 \leq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} + \frac{5K}{n},
-	\end{equation}
-	with probability at least $1 - \exp(-K)$. 
+	\|\wh{f} - f_0\|_n^2 \leq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} + \frac{5K}{n};
+	\end{equation}
+	this is guaranteed if $K = 0$, and otherwise holds with probability at least $1 - \exp(-K)$ if $1 \leq K \leq n$. 
 \end{lemma}
@@
 	\end{equation*}
-	In the above, the last equality relies on the fact that $v_k$ are orthonormal. Using the fact that the eigenvalues are in increasing order, we obtain that
-	\begin{equation*}
-	\sum_{k = K + 1}^n \dotp{v_k}{f_0}_n^2 \leq \frac{1}{\lambda_K^s} \sum_{k = K + 1}^n \lambda_k^s \dotp{v_k}{f_0}_n^2 \leq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s}.
-	\end{equation*}
-
-	The second term in~\eqref{pf:fixed_graph_estimation_1} (estimation error) is random.
+	In the above, the last equality relies on the fact that $v_k$ are orthonormal in $L^2(P_n)$. Using the fact that the eigenvalues are in increasing order, we obtain
+	\begin{equation*}
+	\sum_{k = K + 1}^n \dotp{v_k}{f_0}_n^2 \leq \frac{1}{\lambda_{K + 1}^s} \sum_{k = K + 1}^n \lambda_k^s \dotp{v_k}{f_0}_n^2 \leq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s}.
+	\end{equation*}
+
+	If $K = 0$, $\wh{f} = \Ebb{\wh{f}} = 0$, and the second term in~\eqref{pf:fixed_graph_estimation_1} is $0$.
+Otherwise the second   in~\eqref{pf:fixed_graph_estimation_1} (estimation error) is random.
 Observe that $\dotp{v_k}{\varepsilon}_n \overset{d}{=} Z_k/\sqrt{n}$, where $(Z_1,\ldots,Z_n) \sim N(0,I_{n \times n})$.
@@
 	\end{equation*}
-	Setting $t = K$ completes the proof of the Lemma.
+	Setting $t = K$ completes the proof of the lemma.
 
@@
 \label{subsec:empirical_norm_sobolev_manifold}
-The proof of Proposition~\ref{prop:empirical_norm_sobolev_manifold} follows exactly the same steps as the proof of Proposition~\ref{prop:empirical_norm_sobolev}, except Lemma~\ref{lem:gagliardo_nirenberg} is replaced by Lemma~\ref{lem:gagliardo_nirenberg_manifold}.
+The proof of Proposition~\ref{prop:empirical_norm_sobolev_manifold} follows exactly the same steps as the proof of Proposition~\ref{prop:empirical_norm_sobolev}, upon replacing Lemma~\ref{lem:gagliardo_nirenberg} by Lemma~\ref{lem:gagliardo_nirenberg_manifold}.
 \begin{lemma}[(c.f Theorem~3.70 of~\citet{aubin2012})]
@@
 
-\section{Proofs of Main Results}
+\section{Proof of Main Results}
 \label{sec:proofs_main_results}
@@
 Note that by Propositions~\ref{prop:graph_seminorm_fo} and~\ref{prop:graph_eigenvalue}, these statements are both satisfied with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\}$.
-Conditional on~\eqref{pf:laplacian_eigenmaps_estimation_fo_1}, we have from Lemma~\ref{lem:fixed_graph_estimation} that for any $1 \leq K \leq n - 1$, with probability at least $1 - \exp(-K)$,
-\begin{equation*}
-\|\wh{f} - f_0\|_n^2 \leq C\biggl\{\frac{M^2}{\delta (\lambda_{K + 1}(\Delta_P) \wedge \varepsilon^{-2})} + \frac{K}{n}\biggr\}.
-\end{equation*}
-Further, from the bounds $\varepsilon \leq c_0 K^{-1/d}$ (Assumption~\ref{asmp:parameters_estimation_fo}) and $\lambda_{K + 1}(\Delta_P) \geq c K^{2/d}$ (Weyl's Law) we can simply the above expression the following,
+
+Conditional on~\eqref{pf:laplacian_eigenmaps_estimation_fo_1}, we have from Lemma~\ref{lem:fixed_graph_estimation} that for any $0 \leq K \leq n$,
+\begin{equation*}
+\|\wh{f} - f_0\|_n^2 \leq C\biggl\{\frac{M^2}{\delta (\lambda_{K + 1}(\Delta_P) \wedge \varepsilon^{-2})} + \frac{K}{n}\biggr\},
+\end{equation*}
+either deterministically (when $K = 0$), or with probability at least $1 - \exp(-K)$ (when $K \geq 1$).
+Further, from the bounds $\varepsilon \leq c_0 K^{-1/d}$ (Assumption~\ref{asmp:parameters_estimation_fo}) and $\lambda_{K + 1}(\Delta_P) \geq c (K + 1)^{2/d}$ (Weyl's Law) we can simply the above expression to the following,
 \begin{equation}
 \label{pf:laplacian_eigenmaps_estimation_fo_2}
-\|\wh{f} - f_0\|_n^2 \leq C\biggl\{\frac{M^2}{\delta}K^{-2/d} + \frac{K}{n}\biggr\}.
+\|\wh{f} - f_0\|_n^2 \leq C\biggl\{\frac{M^2}{\delta}(K + 1)^{-2/d} + \frac{K}{n}\biggr\}.
 \end{equation}
 We now upper bound the right hand side of~\eqref{pf:laplacian_eigenmaps_estimation_fo_2}, based on the value of $K$ chosen in~\ref{asmp:parameters_estimation_fo}.
-The classical case is $K = (M^2n)^{d/(2 + d)}$, in which case~\eqref{pf:laplacian_eigenmaps_estimation_fo_2} implies
+When possible we choose $K = \floor{M^2n}^{d/(2 + d)}$ to balance bias and variance, in which case~\eqref{pf:laplacian_eigenmaps_estimation_fo_2} implies
 \begin{equation*}
@@
 \label{subsec:kernel_smoothing}
-In this section we prove Lemma~\ref{lem:kernel_smoothing_laplacian_eigenmaps} (in Section~\ref{subsec:pf_kernel_smoothing_laplacian_eigenmaps}) and Lemma~\ref{lem:kernel_smoothing_bias} (in Section~\ref{subsec:pf_kernel_smoothing_bias}).
-In Section~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we give a sequence $\{p^{(n)}(x), f_0^{(n)}(x)\}_{n \in \mathbb{N}}$ for which the estimator $T_{h,n}\wc{f}$ (kernel extension of Laplacian eigenmaps) dominates direct kernel smoothing of the responses, in the sense that
-\begin{equation*}
-\lim_{n \to \infty} \sup_{h'} \frac{\Ebb \|T_{h_n,n}\wh{f} - f_0\|_P^2}{\Ebb \|T_{h',n}Y - f_0\|_P^2 } = 0.
-\end{equation*}
-
-\subsection{Proof of Lemma~\ref{lem:kernel_smoothing_laplacian_eigenmaps}}
-\label{subsec:pf_kernel_smoothing_laplacian_eigenmaps}
+In this section we prove Lemma~\ref{lem:kernel_smoothing_insample} (in Section~\ref{subsec:pf_kernel_smoothing_insample}) and Lemma~\ref{lem:kernel_smoothing_bias} (in Section~\ref{subsec:pf_kernel_smoothing_bias}).
+In Section~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we give a sequence of design densities and regression functions $\{p^{(n)}(x), f_0^{(n)}(x)\}_{n \in \mathbb{N}}$ for which the estimator $T_{h,n}\wc{f}$ (kernel extension of Laplacian eigenmaps) strictly outperforms directly kernel smoothing the responses, in the sense that
+\begin{equation*}
+\lim_{n \to \infty} \sup_{h'} \frac{\Ebb \|T_{h,n}\wh{f} - f_0\|_P^2}{\Ebb \|T_{h',n}Y - f_0\|_P^2 } = 0.
+\end{equation*}
+
+\subsection{Proof of Lemma~\ref{lem:kernel_smoothing_insample}}
+\label{subsec:pf_kernel_smoothing_insample}
 
@@
 		\end{equation*}
-		\item If $f_0 \in H_0^{s}(P)$, $p \in C^{s}(\mc{X})$, and $\psi$ additionally satisfies~\ref{asmp:order_s_kernel}, it follows that
+		\item If $f_0 \in H_0^{s}(\mc{X})$, $p \in C^{s - 1}(\mc{X})$, it follows that
 		\begin{equation*}
@@
 \label{subsec:upper_bound_l2pn_risk}
-
-\subsection{Facts about manifolds}
 \textcolor{red}{(TODO)}

### 2021-04-29 manuscripts/laplacian_eigenmaps/appendix.tex
Added proof of kernel-smoothing of in-sample estimator.

@@
 
-\textcolor{red}{(TODO)}
+To begin with, we apply the triangle inequality to upper bound $\|T_{h,n}\wc{f} - f_0\|_P$ by the sum of two terms,
+\begin{equation}
+\label{pf:kernel_smoothing_insample_1}
+\|T_{h,n}\wc{f} - f_0\|_P \leq \|T_{h,n}(\wc{f} - f_0)\|_P + \|T_{h,n}f_0 - f_0\|_P.
+\end{equation}
+We proceed by separately upper bounding each term on the right hand side of~\eqref{pf:kernel_smoothing_insample_1}. We will show that
+\begin{equation}
+\label{pf:kernel_smoothing_insample_2}
+\|T_{h,n}(\wc{f} - f_0)\|_P^2 \leq C \|\wc{f} - f_0\|_n^2
+\end{equation}
+and that 
+\begin{equation}
+\label{pf:kernel_smoothing_insample_3}
+\|T_{h,n}f_0 - f_0\|_P^2 \leq \frac{C}{\delta} \cdot \frac{h^2}{nh^d} |f|_{H^1(\mc{X})}^2 + \frac{C}{\delta}\|T_{h,P}f_0 - f_0\|_P^2,
+\end{equation}
+each with probability at least $1 - C\exp(-cnh^d)$. Together these will imply the claim.
+
+\underline{\emph{Proof of~\eqref{pf:kernel_smoothing_insample_2}}.}
+Fix $x \in \mc{X}$. By the Cauchy-Schwarz inequality we have
+\begin{align*}
+\Bigl[T_{h,n}\bigl(\wc{f} - f_0\bigr)(x)\Bigr]^2 & = \Biggl[\frac{1}{d_{n,h}(x)^2}\int \psi\biggl(\frac{\|x' - x\|}{h}\biggr) \cdot \bigl(\wc{f}(x') - f_0(x')\bigr) \,dP_n(x')\Biggr]^2 \\
+& \leq \Biggl[\frac{1}{d_{n,h(x)}^2} \int \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr| \,dP_n(x')\Biggr] \cdot \Biggl[\int \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr| \cdot \bigl(\wc{f}(x') - f_0(x')\bigr)^2 \,dP_n(x')\Biggr] \\
+& = \frac{d_{n,h}^{+}(x)}{|d_{n,h}(x)|} \cdot \frac{1}{|d_{n,h}(x)|} \Biggl[\int \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr| \cdot \bigl(\wc{f}(x') - f_0(x')\bigr)^2 \,dP_n(x')\Biggr].
+\end{align*}
+In the last line all we have done is written $d_{n,h}^{+}(x)$ for the degree functional computed with respect to the kernel $|\psi|$, recalling that $\psi$ may take negative values so $d_{n,h}^{+}(x)$ may not be equal to $d_{n,h}(x)$.
+
+Now we integrate over $x \in \mc{X}$ to get 
+\begin{align}
+\|T_{h,n}(\wc{f} - f_0)\|_P^2 & = \int \Bigl[T_{h,n}\bigl(\wc{f} - f_0\bigr)(x)\Bigr]^2 \,dP(x) \nonumber \\
+& \leq \int \int \frac{d_{n,h}^{+}(x)}{|d_{n,h}(x)|} \cdot \frac{1}{|d_{n,h}(x)|} \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr| \cdot \bigl(\wc{f}(x') - f_0(x')\bigr)^2 \,dP_n(x') \,dP(x) \nonumber \\
+& \leq \sup_{x \in \mc{X}} \frac{d_{n,h}^{+}(x)}{|d_{n,h}(x)|} \cdot \int \int  \frac{1}{|d_{n,h}(x)|} \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr| \cdot \bigl(\wc{f}(x') - f_0(x')\bigr)^2 \,dP(x) \,dP_n(x') \nonumber \\
+& \leq \sup_{x \in \mc{X}} \frac{d_{n,h}^{+}(x) d_{P,h}^+(x)}{|d_{n,h}(x)|^2} \cdot \|\wc{f} - f_0\|_n^2. \label{pf:kernel_smoothing_insample_4}
+\end{align}
+Thus we have reduced the problem to showing that the various degree functionals $d_{n,h}^{+}, d_{P,h}^{+}$ and $d_{n,h}$ all put similar weight on a given point $x$.
+We use~\eqref{eqn:uniform_bound_empirical_degree_2}, which gives a uniform multiplicative bound on deviations of the empirical degree around its mean, to conclude that with probability at least $1 - C\exp\{-cnh^d\}$,
+\begin{equation*}
+d_{n,h}(x) \geq \frac{1}{2}d_{P,h}(x)~~ \textrm{and} ~~ d_{n,h}^{+}(x) \leq \frac{3}{2} d_{P,h}^{+}(x) \quad\textrm{for all $x \in \mc{X}$.}
+\end{equation*}
+Therefore,
+\begin{equation*}
+\sup_{x \in \mc{X}} \frac{d_{n,h}^{+}(x) d_{P,h}^+(x)}{|d_{n,h}(x)|^2} \leq 6 \cdot \sup_{x \in \mc{X}} \frac{|d_{P,h}^{+}(x)|^2}{|d_{P,h}(x)|^2} \leq 36 \biggl(\frac{\|\psi\|_{\infty} p_{\max} \nu_d}{p_{\min}}\biggr)^2
+\end{equation*}
+with the second inequality following from~\eqref{eqn:degree_lower_bound}. Plugging this back into~\eqref{pf:kernel_smoothing_insample_4} gives the claim.
+
+\underline{\emph{Proof of~\eqref{pf:kernel_smoothing_insample_3}}.}
+At a given point $x \in \mc{X}$, we have
+\begin{align*}
+T_{n,h}f_0(x) - f_0(x) & = \frac{1}{d_{n,h}(x)} \sum_{i = 1}^{n} \bigl(f_0(X_i) - f_0(x)\bigr) \psi\biggl(\frac{\|X_i - x\|}{h}\biggr) \\
+& = \frac{d_{P,h}(x)}{d_{n,h}(x)} \cdot \frac{1}{nd_{P,h}(x)} \sum_{i = 1}^{n} \bigl(f_0(X_i) - f_0(x)\bigr) \psi\biggl(\frac{\|X_i - x\|}{h}\biggr).
+\end{align*}
+Thus,
+\begin{equation}
+\label{pf:kernel_smoothing_insample_5}
+\Bigl[T_{n,h}f_0(x) - f_0(x)\Bigr]^2 = \biggl[\frac{d_{P,h}(x)}{d_{n,h}(x)}\biggr]^2 \cdot \biggl[\underbrace{\frac{1}{nd_{P,h}(x)}\sum_{i = 1}^{n} \bigl(f_0(X_i) - f_0(x)\bigr) \psi\biggl(\frac{\|X_i - x\|}{h}\biggr)}_{:= \wt{L}_{n,h}f_0(x)}\biggr]^2
+\end{equation}
+In the proof of~\eqref{pf:kernel_smoothing_insample_2} we have already given an upper bound on the ratio of population to empirical degree, which implies that
+\begin{equation*}
+\sup_{x \in \mc{X}} \biggl[\frac{d_{P,h}(x)}{d_{n,h}(x)}\biggr]^2 \leq 4,
+\end{equation*}
+with probability at least $1 - C\exp\{-cnh^d\}$. On the other hand, we note that the second term in the product in~\eqref{pf:kernel_smoothing_insample_5} has expectation
+\begin{equation*}
+\Ebb\Bigl[\wt{L}_{n,h}f_0(x)\Bigr] = T_{P,h}f_0(x) - f_0(x),
+\end{equation*} 
+and variance 
+\begin{equation*}
+\Var\Bigl[\wt{L}_{n,h}f_0(x)\Bigr] \leq \frac{1}{n(d_{P,h}(x))^2} \Ebb\biggl[(f_0(X) - f_0(x))^2 \cdot \biggl|\psi\biggl(\frac{\|X - x\|}{h}\biggr)\biggr|^2\biggr].
+\end{equation*}
+Integrating with respect to $P$ gives
+\begin{align*}
+\Ebb\biggl[\int \Bigl(\wt{L}_{n,h}f_0(x)\Bigr)^2 \,dP(x)\biggr] & = \int \Ebb\Bigl[\Bigl(\wt{L}_{n,h}f_0(x)\Bigr)^2\Bigr] \,dP(x) \\
+& \leq \| T_{P,h}f_0 - f_0\|_P^2 + \frac{1}{n}\int \int \frac{1}{\bigl(d_{P,h}(x)\bigr)^2} \bigl(f_0(x') - f_0(x)\bigr)^2 \cdot \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr|^2 \,dP(x') \,dP(x) \\
+& \leq \| T_{P,h}f_0 - f_0\|_P^2 + \frac{3}{p_{\min}nh^d} \wt{E}_{P,h}(f_0),
+\end{align*}
+where in the final inequality we have used the lower bound on $d_{P,h}(x)$ from~\eqref{eqn:degree_lower_bound}, and defined $\wt{E}_{P,h}(f_0)$ to be the non-local Dirichlet energy with respect to the kernel $\psi^2$.
+
+Putting the pieces together, we conclude that
+\begin{align*}
+\|T_{n,h}f_0(x) - f_0(x)\|_P^2 & = \int \bigl(T_{n,h}f_0(x) - f_0(x)\bigr)^2 \,dP(x) \\
+& \leq \sup_{x \in \mc{X}} \biggl[\frac{d_{P,h}(x)}{d_{n,h}(x)}\biggr]^2 \cdot \int \Bigl(\wt{L}_{n,h}f_0(x)\Bigr)^2 \,dP(x) \\
+& \overset{(i)}{\leq} 4\frac{\| T_{P,h}f_0 - f_0\|_P^2}{\delta} + \frac{12}{\delta p_{\min}nh^d} \wt{E}_{P,h}(f_0) \\
+& \overset{(ii)}{\leq}  4\frac{\| T_{P,h}f_0 - f_0\|_P^2}{\delta} + \frac{C}{\delta p_{\min}nh^d} h^2 |f_0|_{H^1(\mc{X})}^2
+\end{align*}
+with probability at least $1 - \delta - C\exp(-cnh^d)$.
+In $(i)$ we have used Markov's inequality, and in $(ii)$ we have applied the estimate~\eqref{pf:estimate_nonlocal_seminorm_1} to the non-local Dirichlet energy $\wt{E}_{P,h}(f_0)$.
+This establishes~\eqref{pf:kernel_smoothing_insample_3}.
 
@@
 
-In this section, we give a sequence of densities $p^{(n)}$ and regression functions $f_0^{(n)}$ for which the $L^2(P)$ error of the two-stage estimator $T_{h,n}\wh{f}$ is stochastically dominated by the $L^2(P)$ error of directly kernel smoothing.
+In this section, we give a sequence of densities $p^{(n)}$ and regression functions $f_0^{(n)}$ for which the $L^2(P)$ error of the two-stage estimator $T_{h,n}\wh{f}$ is stochastically dominated by the $L^2(P)$ error of applying kernel smoothing directly to the responses.
 We begin by precisely setting up the problem and reviewing the two estimators we will compare.
@@
 
-\paragraph{A Useful Pair of Lemmas.}
+\paragraph{A Useful Lemma.}
 Let $m = 4n/\log^2n$. Furthermore, for $i = 0,\ldots,m - 1$, let
@@
 
-The following Lemma is from~\cite{laurent00}.
+\paragraph{Proof (of Lemma~\ref{lem:balls_in_bins}).}
+For each $Q_{ij}$, we have that 
+\begin{equation*}
+\Pbb\bigl(\sharp\{Q_{ij} \cup {\bf X}\} = 0\bigr) = (1 - 1/(2m))^{n} \leq \exp\{-n/2m\}.
+\end{equation*}
+Thus by a union bound,
+\begin{equation*}
+\Pbb\Bigl(\sharp\{Q_{ij} \cup {\bf X}\} = 0 ~~\textrm{for any $i = 1,\ldots,m - 1$ and $j = 1,2$} \Bigr) \leq 2m\exp\{-n/2m\} = \frac{8n}{\log^2n}\exp\{-\log^2n/8\},
+\end{equation*}
+and the claim follows from some basic algebra.
+
+\section{Miscellaneous}
+Here we give assorted helpful Lemmas used at various points in the above proofs. We also review notation and relevant facts regarding Taylor expansion.
+
+\subsection{Concentration Inequalities}
+Lemma~\ref{lem:chi_square_bound} controls the deviation of a chi-squared random variable. It is from~\cite{laurent00}.
 \begin{lemma}
@@
 \end{lemma}
-\paragraph{Proof (of Lemma~\ref{lem:balls_in_bins}).}
-For each $Q_{ij}$, we have that 
-\begin{equation*}
-\Pbb\bigl(\sharp\{Q_{ij} \cup {\bf X}\} = 0\bigr) = (1 - 1/(2m))^{n} \leq \exp\{-n/2m\}.
-\end{equation*}
-Thus by a union bound,
-\begin{equation*}
-\Pbb\Bigl(\sharp\{Q_{ij} \cup {\bf X}\} = 0 ~~\textrm{for any $i = 1,\ldots,m - 1$ and $j = 1,2$} \Bigr) \leq 2m\exp\{-n/2m\} = \frac{8n}{\log^2n}\exp\{-\log^2n/8\},
-\end{equation*}
-and the claim follows from some basic algebra.
-
-\section{Miscellaneous}
-Here we give some helpful Lemmas used at various points in the above proofs.
-We also prove that spectral sparsification does not change the rate of convergence of $\wh{f}$.
-Finally, we review notation and relevant facts regarding Taylor expansion.
-
-\subsection{Concentration Inequalities}
-
-The following Lemma is an immediate consequence of the one-sided Bernstein's inequality (14.23) in \cite{wainwright2019}.
+
+Lemma~\ref{lem:one_sided_bernstein} is an immediate consequence of the one-sided Bernstein's inequality (14.23) in \cite{wainwright2019}.
 \begin{lemma}[One-sided Bernstein's inequality]
 	\label{lem:one_sided_bernstein}
-	Let $X, X_1,\ldots,X_n \sim P$, and $f$ satisfy $\Ebb[f^4(X)] < \infty$. Then with probability at least $1 - \exp\bigl(-n/8 \cdot \|f\|_P^4 /\Ebb[f^4(X)]\bigr)$,
-	\begin{equation*}
-	\|f\|_n^2 \geq \frac{1}{2}\|f\|_P^2.
-	\end{equation*}
-\end{lemma}
+	Let $X, X_1,\ldots,X_n \sim P$, and $f$ satisfy $\Ebb[f^4(X)] < \infty$. Then
+	\begin{equation*}
+	\|f\|_n^2 \geq \frac{1}{2}\|f\|_P^2,
+	\end{equation*}
+	with probability at least $1 - \exp\bigl(-n/8 \cdot \|f\|_P^4 /\Ebb[f^4(X)]\bigr)$.
+\end{lemma}
+
+In the proof of Lemma~\ref{lem:kernel_smoothing_insample}, we require uniform control of the empirical degree functional $d_{n,h}(x)$ over all $x \in \mc{X}$.
+Such a result is available to us because the kernel $\psi$ is Lipschitz on its support, so that the class of functions $\{\psi((x - \cdot)/h): h \in \Rd\}$ has finite VC dimension.
+The precise estimate we use is due to \textcolor{blue}{(Gine + Guillou 2002)}.
+\begin{lemma}[Uniform bound for empirical degree.]
+	\label{lem:uniform_bound_empirical_degree}
+	Suppose Model~\ref{def:model_flat_euclidean}.
+For a kernel $\psi$ satisfying~\ref{asmp:kernel} and bandwidth $h$ satisfying~\ref{asmp:bandwidth}, there exist constants $c, C, c_1$ and $C_1$ which do not depend on $h$ or $n$ such that
+	\begin{equation*}
+	\Pbb\biggl(\sup_{x \in \Rd} n \cdot \Bigl|d_{n,h}(x) - d_{P,h}(x)\Bigr| > t\biggr) \leq C\exp\biggl(-c \frac{t^2}{nh^d}\biggr),
+	\end{equation*}
+	for any $t \in \Reals$ satisfying
+	\begin{equation}
+	\label{eqn:uniform_bound_empirical_degree_1}
+	C_1\sqrt{nh^d \log(1/h)} \leq t \leq c_1 nh^d.
+	\end{equation}
+\end{lemma}
+Now we translate Lemma~\ref{lem:uniform_bound_empirical_degree} into a multiplicative bound, which will be more useful for our purposes.
+By assumption, the density $p$ is Lipschitz.
+Letting $L_p$ denote the Lipschitz constant of $p$, we have that
+\begin{align*}
+d_{P,h}(x) & = \int \psi\biggl(\frac{\|x' - x\|}{h}\biggr) p(x') \,dx' \\
+& = h^d \int \psi(\|z\|) p(hz + x) \1\Bigl\{ hz + x \in \mc{X} \Bigr\} \,dz \\
+& \geq h^d p(x) \int \psi(\|z\|) \1\Bigl\{ hz + x  \in \mc{X}\Bigr\} \,dz - L_p h^{d + 1} \|\psi\|_{\infty} \nu_d.
+\end{align*}
+Since by assumption $\mc{X}$ has Lipschitz boundary, setting $c_0$ to be a sufficiently small constant in~\ref{asmp:bandwidth}, we can further deduce that $\int \psi(\|z\|) \1\{hz + x \in \mc{X}\} \,dz \geq 1/3$, and consequently that
+\begin{equation}
+\label{eqn:degree_lower_bound}
+d_{P,h}(x) \geq h^d \frac{p(x)}{3} \geq h^d \frac{p_{\min}}{3} \quad \textrm{for all $x \in \mc{X}$.}
+\end{equation}
+On the other hand, by setting $C_0$ to be a sufficiently large constant in~\ref{asmp:bandwidth}, we can ensure that choosing $t = n h^d p_{\min}/6$ satisfies both the inequalities~\eqref{eqn:uniform_bound_empirical_degree_1}.
+For this choice of $t$ it follows from~\eqref{eqn:degree_lower_bound} and Lemma~\ref{lem:uniform_bound_empirical_degree} that
+\begin{equation}
+\label{eqn:uniform_bound_empirical_degree_2}
+\sup_{x \in \Rd} \biggl|\frac{d_{n,h}(x) - d_{P,h}(x)}{d_{P,h}(x)}\biggr| \leq \sup_{x \in \Rd} \biggl|\frac{d_{n,h}(x) - d_{P,h}(x)}{2t/n}\biggr| \leq \frac{1}{2}
+\end{equation}
+with probability at least $1 - C\exp(-cnh^d)$. This is the form of the result we use in the proof of Lemma~\ref{lem:kernel_smoothing_insample}.
 

### 2021-05-04 notes/misc/splines_graph_laplacians_connections.tex
Partway through updating note to include results on Laplacian smoothing out of sample.

@@
 
-Recall the definition of an \emph{order-$m$ smoothing spline}. For observations $(x_1,y_1),\ldots,(x_n,y_n)$, design points $x_i$ each belonging to a domain $\Omega \subset \Rd$, 
+Recall the definition of an \emph{order-$m$ smoothing spline}.
+For observations $(x_1,y_1),\ldots,(x_n,y_n)$, design points $x_i$ each belonging to a open set with smooth boundary $\Omega \subset \Rd$,
 \begin{equation}
 \label{eqn:smoothing_spline}
-\wh{g} = \argmin_{g \in G} \frac{1}{n} \sum_{i = 1}^{n} (y_i - g(x_i))^2 + \lambda |g|_{m}^2;
+\wh{g} = \argmin_{g \in H^1(\Omega)} \frac{1}{n} \sum_{i = 1}^{n} (y_i - g(x_i))^2 + \lambda |g|_{m}^2;
 \end{equation}
-here we use $|g|_{m}$ to denote the $H^m(\Omega)$, and $G$ to be an appropriately defined function class.
-The standard choice is $G = H^m(\Omega)$, in which case we must also choose $m > d/2$ to make~\eqref{eqn:smoothing_spline} be well-defined.
+here we use $|g|_{m}$ to denote a Hilbert-Sobolev semi-norm on functions $H^m(\Omega)$, and we insist that $m > d/2$ to make~\eqref{eqn:smoothing_spline} be well-defined.
 
-The goal of this note is to review properties of order-$m$ smoothing splines, and see if/when graph Laplacian smoothing possesses analogous properties.
+The goal of this note is to review properties of the order-$m$ smoothing spline, to see if/when graph Laplacian smoothing possesses analogous properties, and to see if/when graph Laplacian smoothing possesses different properties.
 
-\section{Well-posedness of smoothing splines}
-\textcolor{red}{(TODO)}
-
-\section{Optimality of smoothing splines}
-Suppose $x_1,\ldots,x_n$ are sampled independently from a \textcolor{red}{sufficiently regular} density $p$ on $\Omega$, and that
-\begin{equation}
-\label{model:regression}
-y_i = f_0(x_i) + \varepsilon_i,
-\end{equation}
-where $\varepsilon_i \sim N(0,1)$ are i.i.d. Gaussian noise that are also independent of $x_i$. Assume that $f_0 \in H^s(\Omega)$. What can we say about mean squared error $\|\wh{g} - g\|_P^2$?
-
-\textcolor{red}{(TODO)}:
-\begin{itemize}
-	\item Smoothing splines are optimal when $2m > s$. Is there a lower bound on $s$ relative to $m$ or $d$?
-	\item When $m < s < 2s$, the optimality requires some boundary conditions on $f_0$.
-Detail these boundary conditions and explain intuition behind why they are required.
-Compare with the boundary conditions required for orthogonal series/non-parametric least squares estimators to be optimal.
- 	\item Detail various techniques for establishing optimality. Refer to \textcolor{red}{(van der Geer, Nussbaum, Cox)}. Review pros and cons of each technique.
-\end{itemize}
-
-\section{Low-smoothness regime}
-We now introduce a \emph{restricted smoothing spline}, in order to compute estimates for $m < d/2$.
-Suppose we take $\Phi_N \subseteq G$ to be a finite set of $L^2(P)$-orthogonal low-frequency functions; formally $\Phi_N = \{\phi_1,\ldots,\phi_N\}$, where each $\phi_k$ satisfies the Laplacian eigenfunction equation,
-\begin{equation}
-\label{eqn:laplacian_eigen}
-\Delta_P \phi_k = \lambda_k \phi_k,\quad \frac{\partial}{\partial {\bf n}}\phi_k = 0.
-\end{equation}
-Let $F_N = \mathrm{span}\{\Phi_N\}$, and consider
-\begin{equation}
-\label{eqn:restricted_smoothing_spline}
-\wh{g}_{\Phi} = \argmin_{g \in F_N} \frac{1}{n} \sum_{i = 1}^{n} (y_i - g(x_i))^2 + \lambda |g|_{m}^2.
-\end{equation}
-The restricted smoothing spline~\eqref{eqn:smoothing_spline} is well-posed, and its solution $\wh{g}_{\Phi}$ possesses a relatively simple analytic solution.
-It may be totally impossible to exactly compute because it is difficult to exactly solve~\eqref{eqn:laplacian_eigen}, but that is besides the point.
-Now we analyze its statistical properties.
-To do so, we reduce the analysis to that of a related estimator in the Normal means/Gaussian sequence model.
-
-\textcolor{red}{(TODO)}
-\begin{itemize}
-	\item Refer to the results of \textcolor{blue}{(Birge and Massart)} for empirical risk minimization (ERM) over normed balls in H\"{o}lder spaces.
-Resolve why Birge and Massart find a phase transition at $m = 1, d = 2$, which is not at the $4m = d$ boundary; I suspect it is because Birge and Massart find the point at which the empirical process diverges.
-\end{itemize}
-
-\subsection{Gaussian sequence model}
+\section{Gaussian sequence model}
+All calculations are easy in the Gaussian sequence model, and we begin there.
+Much of what we do is extracted from \textcolor{blue}{(Johnstone)}.
 In the Gaussian sequence model, for each $i \in \mathbb{N}$ one observes
@@
 \end{equation}
-Supposes the sequence $\theta^{\ast}$ possesses some ``smoothness''.
-In particular, suppose $\theta^{\ast}$ belongs to the \emph{order-$s$ Sobolev ellipsoid} of radius $M$, $\Theta^s = \{\theta: |\theta|_s \leq M\}$, where for $s \in \mathbb{N}$ the Sobolev seminorm $|\theta|_s$ is defined by
-\begin{equation*}
-|\theta|_s^2 := \sum_{j = 1}^{\infty} \theta_j^2 j^{2s/d}.
-\end{equation*}
-Consider the \emph{sequence smoothing spline} estimator, defined for $N \in \mathbb{N}$ as
+Consider the order-$m$ \emph{sequence smoothing spline} estimator,
 \begin{equation}
 \label{eqn:sequence_smoothing_spline}
-\wh{\theta}_{N} := \argmin_{\theta \in \ell^2(N)} |Y - \theta|_{0} + \lambda |\theta|_m^2, \quad \wh{\theta} := \lim_{N \to \infty} \wh{\theta}_{N}.
+\wh{\theta} := \argmin_{\theta \in \ell^2(\mathbb{N})} |Y - \theta|_{0} + \lambda |\theta|_m^2.
 \end{equation}
-Note that the sequence space estimator $\wh{\theta}$ is perfectly well-posed even though its analogous function space estimator is not. In fact, the sequence space estimator has the explicit solution
+where for $m \in \mathbb{N}$ the Sobolev seminorm $|\theta|_m$ is defined by
+\begin{equation*}
+|\theta|_m^2 := \sum_{j = 1}^{\infty} \theta_j^2 j^{2m/d}.
+\end{equation*}
+Note that the order-$m$ sequence smoothing spline $\wh{\theta}$ is perfectly well-posed when $m < d/2$, even though its analogous function space estimator is not.
+In fact, the sequence space estimator has the explicit solution
 \begin{equation}
@@
 
+Now we analyze the minimax risk of $\wh{\theta}$ over Sobolev ellipsoids. That is, we upper bound the following quantity,
+\begin{equation*}
+\sup_{\theta^{\ast} \in \Theta^s} \Ebb\bigl[|\wh{\theta} - \theta^{\star}|_{0}^2\bigr]
+\end{equation*}
+where $\Theta^s = \{\theta: |\theta|_s \leq M\}$ is the \emph{order-$s$ Sobolev ellipsoid}.
+Note that we do not assume $s = m$.
+Note also that $|\cdot|_0$ is the usual norm over $\ell^2(\mathbb{N})$.
+Finally, a word on terminology: we will use the term ``order-$m$ smoothness'' to refer to the size of the squared norm $|\theta|_m^2$.
+Of course, these are sequences, and the notion of ``smooth'' sequences is not a totally natural concept.
+We use this wording regardless, to unify our discussion of Gaussian sequence and non-parametric regression models.
 \textcolor{red}{(TODO)}
 \begin{itemize}
-	\item Relate the sampling model~\eqref{model:gaussian_sequence_model} to~\eqref{model:regression}, and likewise relate the estimator~\eqref{eqn:sequence_smoothing_spline} to~\eqref{eqn:restricted_smoothing_spline}.
-Refer to \textcolor{blue}{(Tsybakov, Johnstone)}.
-	\item Show that in the Gaussian sequence model, if $m > \max\{d/4,s/2\}$ then~\eqref{eqn:sequence_smoothing_spline} achieves optimal $\ell_2$ loss when $N = \infty$.
+	\item Show that in the Gaussian sequence model, if $m > \max\{d/4,s/2\}$ then~\eqref{eqn:sequence_smoothing_spline} achieves optimal risk.
 \end{itemize}
 \subsection{``Failure'' of restricted smoothing splines in the very-low-smoothness regime}
-On the other hand, whenever $m \leq d/4$ meaningful restrictions on the size of $N$ kick in.
+On the other hand, whenever $m \leq d/4$ the estimator $\wh{\theta}$ is grossly sub-optimal. 
 \begin{lemma}
@@
 \begin{itemize}
-	\item When $r = 1$---corresponding to the first-order Sobolev norm--- Lemma~\ref{lem:smoothness_of_sequence_smoothing_splines} tells us that the sequence smoothing spline $\wh{\theta}$ has infinite $|\cdot|_1$ norm when $d \geq 2$.
-Remarkably, when $d = 2$ or $d = 3$, $\wh{\theta}$ continues to be optimal despite being an incredibly ``wiggly'' estimator of a ``smooth'' sequence.
-Note that this forecloses upon the possibility of analyzing $\wh{\theta}$ using uniform convergence over $\Theta^1(M)$, since $\wh{\theta}$ does not belong to this class for any finite $M$.
-	\item When $r = 0$---corresponding to the zeroth-order Sobolev norm, or in other words the $\ell^2$-norm---Lemma~\ref{lem:smoothness_of_sequence_smoothing_splines} tells us that the sequence smoothing spline $\wh{\theta}$ has infinite $\Theta^0$ norm when $d \geq 4$.
+	\item Consider $r = 1$, which corresponds to the first-order Sobolev norm.
+Lemma~\ref{lem:smoothness_of_sequence_smoothing_splines} tells us that the sequence smoothing spline $\wh{\theta}$ has infinite first-order Sobolev norm in expectation, $\Ebb|\wh{\theta}|_1 = \infty$, whenever $d \geq 2$.
+Remarkably, when $d = 2$ or $d = 3$, $\wh{\theta}$ continues to have optimal risk despite being an incredibly ``wiggly'' estimator of a ``smooth'' sequence.
+Note that this forecloses on the possibility of analyzing $\wh{\theta}$ using uniform convergence over $\Theta^1(M)$, since $\wh{\theta}$ does not belong to this class for any finite $M$.
+	\item Consider $r = 0$, which corresponds to the zeroth-order Sobolev norm, or in other words the $\ell^2$-norm.
+Lemma~\ref{lem:smoothness_of_sequence_smoothing_splines} tells us that the sequence smoothing spline $\wh{\theta}$ has infinite $\Theta^0$ norm when $d \geq 4$.
 Clearly, since the regression function $f_0$ has finite $\Theta^0$ norm, and the estimator $\wh{f}$ has infinite $\Theta^0$ norm, it follows from the triangle inequality that $|\wh{f} - f|_0 = \infty$.
@@
 \begin{itemize}
+	\item Do an analogous calculation for empirical risk minimization over $\Theta^1(M)$.
+Intuition, both from the above calculations and from \textcolor{blue}{(Birge and Massart)}, suggest that there is a phase transition at $d = 2$.
 	\item Do an analogous calculation for the $\Theta^{1,1}$ case, with the estimator
@@
 
-\section{Relationship to Graph Laplacian methods}
+\section{High-smoothness regime: optimality of smoothing splines}
+Suppose $x_1,\ldots,x_n$ are~\textcolor{red}{(well-spaced)} in $\Omega$, and that
+\begin{equation}
+\label{model:regression}
+y_i = f_0(x_i) + \varepsilon_i,
+\end{equation}
+where $\varepsilon_i \sim N(0,1)$ are i.i.d.
+Gaussian noise that are also independent of $x_i$.
+Assume that $f_0 \in H^s(\Omega)$.
+What can we say about in-sample mean squared error $\|\wh{g} - g\|_n^2$?
+What can we say about the out-of-sample mean squared error $\|\wh{g} - g\|_{L^2(\mc{X})}^2$?
+
+\textcolor{red}{(TODO)}:
+\begin{itemize}
+	\item Necessary conditions for the order-$m$ smoothing spline to be optimal are that $m > s/2$ and $m > d/2$.
+Are these sufficient conditions---at least, on the indices $(m,s,d)$---or is there a lower bound on $s$ relative to $m$ or $d$?
+	\item When $s/2 < m < s$, the aforementioned optimality properties require that $f_0$ satisfy some boundary conditions.
+Detail these boundary conditions and explain why they are necessary.
+Compare with the boundary conditions required for orthogonal series/non-parametric least squares estimators to be optimal.
+ 	\item Detail various techniques for establishing optimality. Refer to \textcolor{red}{(van der Geer, Nussbaum, Cox)}. Review pros and cons of each technique.
+\end{itemize}
+
+\section{Low-smoothness regime: optimality of restricted smoothing splines}
+We now introduce the order-$m$ \emph{restricted smoothing spline}, an alternative to the order-$m$ smoothing spline which makes sense when $m < d/2$.
+To define the order-m restricted smoothing spline, we first need to review the \emph{weighted Laplace-Beltrami operator} $\Delta_P$, and an associated eigenvector problem.
+For this purpose, we now introduce our formal assumptions on the design distribution $P$.
+Let $\nu$ denote the $d$-dimensional Lebesgue measure.
+\begin{assumption}[Design distribution $P$.]
+	\label{asmp:design}
+	The design distribution $P$ is supported on an open, bounded set $\Omega \subseteq \Rd$, which has $C^{\infty}$ boundary $\partial \Omega$.
+The distribution $P$ admits a $C^{\infty}(\Omega)$ density $p$ with respect to Lebesgue measure $\nu$, which is bounded away from $0$ and $\infty$, meaning that for some $p_{\min} < 1 < p_{\max}$
+	\begin{equation*}
+	0 \leq p_{\min} \leq p(x) \leq p_{\max} < \infty,\quad\textrm{for all $x \in \Omega$.}
+	\end{equation*}
+\end{assumption} 
+We note that the assumptions in~\eqref{asmp:design} are in a certain sense very strong, and quite likely overkill for any particular value of $s$ and $d$.
+The minimum conditions required on $P$ in order to have the operator $\Delta_P$ and its eigenvectors be ``sufficiently regular'' for statistical purposes is certainly of interest, but not the point of this note.
+
+If $P$ satisfies Assumption~\ref{asmp:design}, the weighted Laplace-Beltrami operator $\Delta_P$ is defined as 
+\begin{equation*}
+\Delta_Pf := \frac{-1}{p} \mathrm{div}(p^2 \nabla f)~~\textrm{in $\Omega$}, \quad \frac{\partial f}{\partial {\bf n}} = 0~~\textrm{on $\partial \Omega$.}
+\end{equation*}
+(The second part of the statement above should interpreted in the usual way, to mean that $\Delta_P$ is an operator acting on those functions $f \in L^2(P)$ that satisfy Neumann boundary conditions).
+We see that $\Delta_P$ is a second-order symmetric differential operator, and is self-adjoint in $L^2(P)$; the assumptions on $p$ further imply that it is uniformly elliptic.
+As such it admits a discrete spectrum, with eigenvalues $0 < \lambda_1 \leq \lambda_2 \leq \cdots$, and corresponding eigenvectors $\psi_1,\psi_2,\ldots$ which form an orthonormal basis of $L^2(P)$.
+Standard theory (e.g.
+Theorem 6, Chapter 6.3 of \textcolor{blue}{(Evans)}) implies that each eigenvector $\psi_j \in C^{\infty}(\wb{\Omega})$.
+
+Let $N$ be a positive integer.
+We define the order-$m$ restricted smoothing spline to be the penalized least squares estimator over the linear subspace  $\Phi_N = \mathrm{span}\{\psi_j: j \leq N\}$ of $L^2(P)$.
+Formally speaking,
+\begin{equation}
+\label{eqn:restricted_smoothing_spline}
+\wh{g}_{N} = \argmin_{g \in \Phi_N} \frac{1}{n} \sum_{i = 1}^{n} (y_i - g(x_i))^2 + \lambda |g|_{m}^2.
+\end{equation}
+The order-$m$ restricted smoothing spline defined in~\eqref{eqn:restricted_smoothing_spline} is well-posed.
+Let $\Gamma \in \Reals^{N \times N}$ be the matrix with entries $\Gamma_{ij} = \dotp{\psi_i}{\psi_j}_{m}$.
+For a given $\alpha \in \Reals^N$, define $g_{\alpha} = \sum_{i = j}^{N} \alpha_j \psi_j$.
+Then
+\begin{equation*}
+\wh{g}_{N} = \argmin_{\alpha \in \Reals^n} \frac{1}{n} \sum_{i = 1}^{n} (y_i - g_{\alpha}(x_i))^2 + \lambda \alpha^T \Gamma \alpha.
+\end{equation*}
+This is just a generalized ridge problem, and can be solved either exactly, or approximately by an iterative procedure.
+
+Unfortunately, it may be totally impossible to actually compute $\wh{g}_N$, because in order to do so we need access to the eigenfunctions $\psi_j$, which in turn requires knowing $P$.
+We ignore this problem for the moment, but return to it later when we introduce graph Laplacian smoothing methods.
+For now, we assume that we can compute the estimator defined by~\eqref{eqn:restricted_smoothing_spline}, and concentrate on analyzing its statistical properties.
+
+\textcolor{red}{(TODO)}
+\begin{itemize}
+	\item Analyze the statistical properties of the estimator defined in~\eqref{eqn:restricted_smoothing_spline}.
+Note that you are no longer clear about whether or not these are the same as the properties of the sequence model.
+\end{itemize}
+
+\section{Supervised graph Laplacian smoothing}
 
@@
 
+\section{Semi-supervised graph Laplacian smoothing}
+
+Suppose now that in addition to the \emph{labeled} data $(x_1,y_1),\ldots,(x_n,y_n)$, we observe $m$ \emph{unlabeled} design points $x_{n + 1},\ldots,x_{n + m}$.
+Let $N = n + m$, and let $L_{N,\varepsilon}$ denote the graph Laplacian over labeled and unlabeled points,
+\begin{equation*}
+\bigl(L_{N,\varepsilon}u\bigr)_i := \frac{1}{N \varepsilon^{d + 2}}\sum_{j = 1}^{N} (u_i - u_j) \eta\biggl(\frac{\|x_i - x_j\|}{\varepsilon}\biggr).
+\end{equation*}
+The order-$m$ graph Laplacian smoothing estimator $\wh{f}$ defined in \textcolor{red}{(?)} can be naturally generalized to a vector in $\Reals^N$ according to the following formulation,
+\begin{equation}
+\label{eqn:ssl_graph_laplacian_smoothing_0}
+\wh{f}_N := \argmin_{f \in \Reals^N} \|Y - f\|_n^2 + \lambda \dotp{L_{N,\varepsilon}f}{f}_N.
+\end{equation}
+The estimator $\wh{f}$ has a closed-form solution, which can be simply written using the extension operator $E_{n \to N}: \Reals^{n} \to \Reals^N$ and its adjoint, the restriction operator $R_{N \to n}: \Reals^{N} \to \Reals^n$, which are defined by
+\begin{equation*}
+E_{n \to N}u := (u_1,\ldots,u_n,0,\ldots,0),\quad\textrm{and}\quad R_{N \to n}v := (v_1,\ldots,v_n).
+\end{equation*}
+Then the order-$m$ graph Laplacian smoothing estimator can be written as follows
+\begin{equation}
+\label{eqn:ssl_graph_laplacian_smoothing_1}
+\wh{f}_N = (E_{n \to N} R_{N \to n} + \frac{\rho n}{N} L_{N,\varepsilon})^{-1} E_{n \to N}Y.
+\end{equation}
+The form of~\eqref{eqn:ssl_graph_laplacian_smoothing_1} is a compact way to represent the solution $\wh{f}$, but it is rather opaque.
+In Proposition~\ref{prop:ssl_graph_laplacian_smoothing} we give a second form of $\wh{f}$, which is less compact but more evocative.
+\begin{proposition}
+	\label{prop:ssl_graph_laplacian_smoothing}
+	Let $\wh{f}_N$ be the solution to~\eqref{eqn:ssl_graph_laplacian_smoothing_0}. Writing $\wh{f}_N = (\wh{f}_n,\wh{f}_m)$ for $\wh{f}_n \in \Reals^m$ and $\wh{f}_n \in \Reals^m$, we have the following.
+	\begin{itemize}
+		\item 
+	\end{itemize}
+\end{proposition}
+
 \end{document}

### 2021-05-13 code/configs/estimation_wigglycosine1d.R
Updated code base.

@@ -1,5 +1,5 @@
 #------------------------------------------------#
-# Configs for first estimation MSE experiment.
+# Configs for poster presentation experiment.
 #------------------------------------------------#
 
 ### General configs.

### 2021-05-13 code/configs/laplacian_eigenmaps/eigenmaps_approximates_spectral_projection.R

@@ -1,0 +1,20 @@
+#------------------------------------------------#
+# Configs for Laplacian Eigenmaps Figure 1.
+#------------------------------------------------#
+
+### General configs.
+d <- 1
+# ns <- 2000
+ns <- round(seq(400,2000,length.out = 10))
+iters <- 50
+
+### Configs for sampling. ###
+sample_X <- make_sample_uniform(d)
+make_f0 <- make_spectral_sobolev
+
+
+### Methods. ###
+methods <- list(laplacian_eigenmaps = make_laplacian_eigenmaps,
+                spectral_projection = make_spectral_projection)
+initialize_thetas <- list(initialize_laplacian_eigenmaps_thetas,
+                          initialize_spectral_projection_thetas)

### 2021-05-13 code/configs/laplacian_eigenmaps/eigenmaps_beats_least_squares.R

@@ -1,0 +1,22 @@
+#------------------------------------------------#
+# Configs for Laplacian Eigenmaps Figure 2.
+#------------------------------------------------#
+
+### General configs.
+d <- 1
+ns <- round(seq(400,2000,length.out = 10))
+iters <- 50
+
+### Configs for sampling. ###
+sample_X <- make_sample_gaussian_mixture(d,mu = c(-1,1),pi = c(3/4, 1/4), sigma = c(.4,.4))
+# sample_X <- make_sample_uniform(d)
+make_f0 <- make_doppler
+
+
+### Methods. ###
+methods <- list(laplacian_eigenmaps = make_laplacian_eigenmaps,
+                least_squares = make_least_squares,
+                kernel_smoothing = make_kernel_smoothing)
+initialize_thetas <- list(initialize_laplacian_eigenmaps_thetas,
+                          initialize_spectral_projection_thetas,
+                          initialize_kernel_smoothing_thetas)

### 2021-05-13 code/estimation_mse_pipeline.R

@@ -2,12 +2,16 @@ library(RANN)
 library(dplyr)
 library(Matrix)
 library(reshape2)
+library(RSpectra)
 source("sample.R")
 source("graph.R")
 source("estimators.R")
 
 # Please change this line to whichever config you wish to run.
-source("configs/estimation_wigglycosine1d.R")
+source("configs/laplacian_eigenmaps/eigenmaps_approximates_spectral_projection.R")
+
+# Please change this to dictate how verbose the pipeline should be 
+verbose <- F
 
 #----------------------------------------------------#
 # This is the pipeline for running regression estimation experiments.
@@ -20,12 +24,14 @@ Xs <- vector(mode = "list", length = length(ns))
 f0s <- vector(mode = "list", length = length(ns))
 Ys <- vector(mode = "list", length = length(ns))
 thetas <- vector(mode = "list", length = length(ns))
+fits <- vector(mode = "list", length = length(ns))
 
 for(ii in 1:length(ns))
 {
   n <- ns[ii]
   f0 <- make_f0(d,n)
   mse[[ii]] <- vector(mode = "list", length = length(methods))
+  fits[[ii]] <- vector(mode = "list",length = length(methods))
 
   # Initialize tuning parameters (thetas).
   # Note: I have granted methods the right to access the distribution P from which X is sampled
@@ -35,6 +41,7 @@ for(ii in 1:length(ns))
   {
     thetas[[ii]][[jj]] <- initialize_thetas[[jj]](sample_X, n)
     mse[[ii]][[jj]] <- matrix(nrow = nrow(thetas[[ii]][[jj]]), ncol = iters)
+    fits[[ii]][[jj]] <- array(dim = c(nrow(thetas[[ii]][[jj]]),n,iters))
   }
   for(iter in 1:iters)
   {
@@ -50,17 +57,30 @@ for(ii in 1:length(ns))
     {
       method <- methods[[jj]]
       fits_method <- matrix(nrow = nrow(thetas[[ii]][[jj]]), ncol = n)
+      
+      # Do any computations here that we would have to repeat for each hyperparameter (theta)
+      # E.g. compute eigenvectors.
+      precompute_data <- attributes(method)$precompute_data
+      if(!is.null(precompute_data)){
+        precomputed_data <- precompute_data(Y,X, thetas[[ii]][[jj]])
+      } else{
+        precomputed_data <- NULL
+      }
       for(kk in 1:nrow(thetas[[ii]][[jj]])){
[…]

### 2021-05-13 code/estimators.R

@@ -1,3 +1,55 @@
+make_laplacian_eigenmaps <- function(theta)
+{
+  r <- theta[["r"]]
+  K <- theta[["K"]]
+  laplacian_eigenmaps <- function(Y,X){
+    # If we've already computed the eigenvectors don't do it again
+    if(exists("precomputed_data")){
+      stopifnot(r %in% names(precomputed_data))
+      L_eigenvectors <- precomputed_data[[which(names(precomputed_data) == r)]]
+    } else{
+      # Build G_{n,r} over X.
+      G <- neighborhood_graph(X,r)
+      
+      # Get L.
+      L <- Laplacian(G)
+      
+      # Get spectra.
+      spectra <- get_spectra(L,K)
+      L_eigenvectors <- spectra$vectors[,ncol(spectra$vectors):1]
+    }
+    V_K <- L_eigenvectors[,1:K] # Eigenvectors we need
+    a_k <- t(V_K) %*% Y         # Empirical Fourier coefficients
+    f_hat <- V_K %*% a_k        # Spectral projection
+    
+    # browser()
+    
+    return(f_hat)
+  }
+}
+attr(make_laplacian_eigenmaps,"precompute_data") <- function(Y,X, theta_df){
+  rs <- unique(theta_df$r)
+  precomputed_data <- vector(mode = "list",length = length(rs))
+  names(precomputed_data) <- rs
+  for(kk in 1:length(rs)){
+    r <- rs[kk]
+    K <- thetas[[ii]][[jj]] %>% filter(r == !!r) %>% pull(K)
+    
+    # Build G_{n,r} over X.
+    G <- neighborhood_graph(X,r)
+    
+    # Get L.
+    L <- Laplacian(G)
+    
+    # Compute as many eigenvectors as we will need.
+    spectra <- get_spectra(L,max(K))
+    L_eigenvectors <- spectra$vectors[,ncol(spectra$vectors):1]
+    
+    precomputed_data[[kk]] <- L_eigenvectors
+  }
+  return(precomputed_data)
+}
+
 make_laplacian_smoothing <- function(theta)
 {
   r <- theta[["r"]]
@@ -34,6 +86,35 @@ make_laplacian_smoothing_knn <- function(theta)
   }
 }
 
+make_spectral_projection <- function(theta)
+{
+  K <- theta[["K"]]
+  spectral_projection <- function(Y,X)
+  {
+    d <- ncol(X)
[…]

### 2021-05-13 code/graph.R

@@ -38,7 +38,7 @@ knn_to_neighborhood_graph <- function(X, r, k_max)
   {
     A <- Matrix(0, nrow = n, ncol = n)
     k_max <- min(2*k_max,n - 1)
-    rneighbors <- nn2(data = X, query = X, searchtype = 'radius', k = k_max, radius = epsilon)$nn.idx[,-1]
+    rneighbors <- nn2(data = X, query = X, searchtype = 'radius', k = k_max, radius = r)$nn.idx[,-1]
     rneighbors[rneighbors == 0] <- NA
     r_neighbors_list <- as.matrix(melt(rneighbors)[,-2])
     r_neighbors_list <- r_neighbors_list[rowSums(is.na(r_neighbors_list)) == 0,]
@@ -81,4 +81,20 @@ Laplacian <- function(A)
     L <- as(D - A,"dgCMatrix")
   }
   return(L)
+}
+
+# A wrapper around eigs_sym, which takes care of some failure cases when the
+# matrix is poorly conditioned. 
+get_spectra <- function(A,K,sigma = 0){
+  # Compute as many eigenvectors as we will need.
+  spectra <- tryCatch(eigs_sym(A,K,sigma),
+                      error = function(e){
+                        test <- eigs_sym(A,K, which = "SM")
+                        return(test)
+                      })
+  if(max(abs(spectra$values)) > 2e10)
+  {
+    spectra <- eigs_sym(A,K,which = "SM")
+  }
+  spectra
 }

### 2021-05-13 code/plot.R

@@ -3,7 +3,7 @@ library(ggplot2)
 library(gridExtra)
 
 ### Plot. ###
-save_directory <- "data/AISTATS/cosine/d5_largesample" # Please change this to whichever directory you prefer.
+save_directory <- "data/20210510132401" # Please change this to whichever directory you prefer.
 plot_directory <- file.path(save_directory,"plots")
 dir.create(plot_directory)
 
@@ -15,7 +15,7 @@ load(file.path(save_directory,"thetas.R"))
 load(file.path(save_directory,"mse.R"))
 load(file.path(save_directory,"Xs.R"))
 load(file.path(save_directory,"f0s.R"))
-# load(file.path(save_directory,"Ys.R"))
+load(file.path(save_directory,"Ys.R"))
 
 # Plot of mse by tuning parameter.
 pdf(file.path(plot_directory,"mse.pdf"))
@@ -29,6 +29,16 @@ for(jj in 1:length(methods))
     mean_mse <- rowMeans(mse_ii_jj)
     sd_mse <- apply(mse_ii_jj,1,sd)
     
+    if(names(methods)[jj] == "laplacian_eigenmaps"){
+      plot_df <- data.frame(x = thetas_ii_jj[,2],y = mean_mse,
+                            col = as.factor(round(thetas_ii_jj[,1],3)))
+      tuning_plots[[ii]] <- 
+        ggplot(data = plot_df,aes(x = x,y = y, color = col)) + geom_line() + 
+        labs(x = "K",
+             y = "mse",
+             title = paste0("n = ",ns[ii],"")) +
+        theme_bw()
+    }
     if(names(methods)[jj] == "laplacian_smoothing" || is.null(methods))
     {
       plot_df <- data.frame(x = log(thetas_ii_jj[,2]),y = mean_mse,
@@ -94,11 +104,12 @@ if(d == 1)
       plot_name <- paste0(names(methods)[jj],"_estimate_",ii,".png")
       png(file.path(plot_directory,plot_name))
       plot_df <- data.frame(x = Xs[[ii]], y = Ys[[ii]],fhat = best_fits_by_method[[ii]][[jj]])
+      title <- names(methods)[jj]
       plot(x = plot_df$x,y = plot_df$y, 
            col = col,
[…]

### 2021-05-13 code/plot_demos/eigenmaps_figure_1.R

@@ -1,0 +1,155 @@
+library(dplyr)
+library(ggplot2)
+library(gridExtra)
+library(viridis)
+source("plot_methods.R")
+
+# User entered information.
+data_directory <- "data/laplacian_eigenmaps/20210511125310"
+plot_directory <- "plots/laplacian_eigenmaps/eigenmaps_approximates_spectral_projection" # Please change this to whichever directory you prefer.
+if(!exists(plot_directory)){dir.create(plot_directory)}
+
+# Load data
+load(file.path(data_directory,"configs.R"))
+d <- configs$d; ns <- configs$ns; methods <- configs$methods
+load(file.path(data_directory,"best_fits_by_method.R"))
+load(file.path(data_directory,"thetas.R"))
+load(file.path(data_directory,"mse.R"))
+load(file.path(data_directory,"Xs.R"))
+load(file.path(data_directory,"f0s.R"))
+load(file.path(data_directory,"Ys.R"))
+
+# Plotting parameters for all plots
+
+## Plot 1: Regression fxn and fitted values.
+plot1_indx <- 5
+plot1_K <- 10
+rs <- unique(thetas[[plot1_indx]][[1]]$r)
+plot1_r <- rs[which.min(abs(rs - .02))]    # arbitrary choice
+
+# Regression fxn
+pdf(file.path(plot_directory,"f0.pdf"))
+par(mar = c(5.1,6,4.1,4.1)) # Prevent the left hand side from getting cut off.
+plot_fxn(Xs[[plot1_indx]],Ys[[plot1_indx]],f = f0s[[plot1_indx]])
+dev.off()
+
+# Eigenmaps fitted values
+pdf(file.path(plot_directory,"eigenmaps.pdf"))
+par(mar = c(5.1,6,4.1,4.1)) # Prevent the left hand side from getting cut off.
+f <- subset(fits[[plot1_indx]][[1]], 
+            thetas[[plot1_indx]][[1]]$r == plot1_r & 
+            thetas[[plot1_indx]][[1]]$K == plot1_K)[1,]
+            
+plot_fxn(Xs[[plot1_indx]],Ys[[plot1_indx]],f,title = "Eigenmaps, K = 10",col = "red")
+dev.off()
+
+# Eigenmaps fitted values
[…]

### 2021-05-13 code/plot_methods.R

@@ -1,0 +1,90 @@
+plot_fxn <- function(x,y,f,title = NULL,col = NULL){
+    if(is.null(col)){col <- "black"}
+    plot_df <- data.frame(x = x, y = y,f = f)
+    plot(x = plot_df$x,y = plot_df$y, 
+         col = "grey34",cex = .6,
+         cex.main = 2.5, cex.lab = 2.5, cex.axis = 2,
+         lwd = 1.1,xlab = "",ylab = "", main = title)
+    lines(x = x[order(x)],y = f[order(x)],lwd = 1.5,
+          col = col)
+}
+
+
+# Plot of mse---for best choice of tuning parameter---by sample size
+plot_best_mse <- function(methods,mse){
+  cols <- c("red","blue","green")[1:length(methods)]     
+  
+  plot_dfs_best_mse <- vector(mode = "list", length = length(methods))
+  names(plot_dfs_best_mse) <- names(methods)
+  fitted_slopes <- numeric(length(methods))
+  names(fitted_slopes) <- names(methods)
+  for(jj in 1:length(methods))
+  {
+    best_mse <- numeric()
+    minimax_mse <- numeric()
+    sd_best_mse <- numeric()
+    for(ii in 1:length(ns))
+    {
+      mse_ii_jj <- mse[[ii]][[jj]]
+      best_mse[ii] <- min(rowMeans(mse_ii_jj))
+      sd_best_mse[ii] <- apply(mse_ii_jj,1,sd)[which.min(rowMeans(mse_ii_jj))]/sqrt(ncol(mse_ii_jj))
+      minimax_mse[ii] <- ns[ii]^{-2/(2 + d)}
+    }
+    
+    # Rescale minimax mse to match intercept with best_mse
+    minimax_mse <- minimax_mse * (best_mse[1]/minimax_mse[1])
+    plot_dfs_best_mse[[jj]] <- data.frame(x = ns, y = best_mse,sd = sd_best_mse) 
+    
+    # fitted slope
+    log_best_mse <- log(best_mse)
+    log_ns <- log(ns)
+    fitted_slopes[jj] <- round( lm(log_best_mse ~ log_ns)$coefficients[2], 2)
+    
+    # hack to change names for plotting
+    name <- names(plot_dfs_best_mse)[[jj]]
+    names(plot_dfs_best_mse)[[jj]] <- case_when(
+      name == "laplacian_smoothing" ~ "LS",
[…]

### 2021-05-13 code/sample.R

@@ -1,9 +1,25 @@
-make_sample_gaussian <- function(d)
+make_sample_gaussian <- function(d, sigma = 1)
 {
   sample_X <- function(n){
     X <- matrix(NA,nrow = n,ncol = d)
     while(any(is.na(X))){
-      X[which(is.na(X))] <- rnorm(sum(is.na(X)))
+      X[which(is.na(X))] <- rnorm(sum(is.na(X)),0,sigma)
+      X[abs(X) > 1] <- NA
+    }
+    return(X)
+  }
+}
+
+make_sample_gaussian_mixture <- function(d,mu = NULL,pi = NULL,sigma = NULL)
+{
+  if(is.null(mu)){mu <- 0}
+  if(is.null(pi)){pi <- rep(1,length(mu))}
+  if(is.null(sigma)){sigma <- rep(1,length(mu))}
+  sample_X <- function(n){
+    X <- matrix(NA,nrow = n,ncol = d)
+    while(any(is.na(X))){
+      class <- apply(rmultinom(sum(is.na(X)), 1, prob = pi),2,FUN = function(i){which(i != 0)})
+      X[which(is.na(X))] <- sapply(class,FUN = function(l){rnorm(1,mu[l],sigma[l])})
       X[abs(X) > 1] <- NA
     }
     return(X)
@@ -18,10 +34,38 @@ make_sample_uniform <- function(d)
   }
 }
 
+#-----------------------------------------------#
+# Regression functions
+#-----------------------------------------------#
+
+make_doppler <- function(d,n){
+  stopifnot(d == 1)
+  f0 <- function(x){
+    if(x > 0){cos(4*pi/((abs(x)^{1/3})))}
+    else{cos(pi/(abs(x)^{1/3}))}
+  }
+}
+
+make_spectral_sobolev <- function(d,n){
+  #------------------------------#
+  # Fixes a vector beta that belongs to a Sobolev ellipsoid,
+  # then builds the regression function
+  # 
+  # f0(x) = sum_{k} beta_k psi_k
+  #
+  # where psi_k(x) = cos(k*pi*x)
+  stopifnot(d == 1) # Only do 1d for the moment
+  c <- sqrt(sum((1:3)^2))
+  beta <- c(rep(1/c,3),1/(sqrt(17)*(4:20)))
+  f0 <- function(x){
[…]

### 2021-05-13 notes/misc/splines_graph_laplacians_connections.tex
Updated document on connections between splines and graph Laplacians, with a section explaining how Laplacian regularization in unlabeled data case performs out-of-sample. This should probably be put into a different document actually.

@@
 \label{eqn:smoothing_spline}
-\wh{g} = \argmin_{g \in H^1(\Omega)} \frac{1}{n} \sum_{i = 1}^{n} (y_i - g(x_i))^2 + \lambda |g|_{m}^2;
+\wh{g} = \argmin_{g \in H^m(\Omega)} \frac{1}{n} \sum_{i = 1}^{n} (y_i - g(x_i))^2 + \lambda |g|_{m}^2;
 \end{equation}
@@
 \section{Gaussian sequence model}
-All calculations are easy in the Gaussian sequence model, and we begin there.
+All calculations are easier in the Gaussian sequence model, and we begin there.
 Much of what we do is extracted from \textcolor{blue}{(Johnstone)}.
@@
 Of course, these are sequences, and the notion of ``smooth'' sequences is not a totally natural concept.
-We use this wording regardless, to unify our discussion of Gaussian sequence and non-parametric regression models.
+We use this wording regardless, to unify our discussion of the Gaussian sequence and non-parametric regression models.
 \textcolor{red}{(TODO)}
@@
 	\label{lem:failure_very_low_smoothness}
-	Suppose we observe data $Y_1,Y_2,\ldots$ according to Model~\eqref{model:gaussian_sequence_model}. If we compute $\wh{\theta}$ and any $m \leq d/4$, then
+	Suppose we observe data $Y_1,Y_2,\ldots$ according to Model~\eqref{model:gaussian_sequence_model}. If we compute $\wh{\theta}$ as in~\eqref{eqn:sequence_smoothing_spline} with $m \leq d/4$, then
 	\begin{equation}
@@
 	Confirm your expectation: that when error measured in $|\cdot|_0$ norm and $d > 1$, this estimator is optimal when $N = n$, but not when $N = \infty$.
-This would give some evidence that the graph is saving the TV denoiser when $d > 1$.
+This would give some evidence that the graph is saving the TV denoiser when $d > 1$; although to be clear the TV denoiser has no exact sequence model analogue.
 	\item Figure out other smoothness properties of $\wh{\theta}_N$, for instance ``H\"{o}lder'' smoothness.
@@
 What can we say about in-sample mean squared error $\|\wh{g} - g\|_n^2$?
-What can we say about the out-of-sample mean squared error $\|\wh{g} - g\|_{L^2(\mc{X})}^2$?
+What can we say about the out-of-sample mean squared error $\|\wh{g} - g\|_{L^2(\Omega)}^2$?
 
@@
 To define the order-m restricted smoothing spline, we first need to review the \emph{weighted Laplace-Beltrami operator} $\Delta_P$, and an associated eigenvector problem.
-For this purpose, we now introduce our formal assumptions on the design distribution $P$.
+For this purpose, we now introduce some assumptions on the design distribution $P$.
 Let $\nu$ denote the $d$-dimensional Lebesgue measure.
@@
 	\item Analyze the statistical properties of the estimator defined in~\eqref{eqn:restricted_smoothing_spline}.
-Note that you are no longer clear about whether or not these are the same as the properties of the sequence model.
+Note that you are no longer clear about whether or not these are the same as the properties of the sequence model, especially with regards to out-of-sample loss.
 \end{itemize}
@@
 \begin{itemize}
-	\item Review why your analysis of Laplacian smoothing breaks down when $d > 4$, making reference to the previous section.
+	\item Review why your analysis of graph Laplacian smoothing breaks down when $d > 4$, making reference to the previous section.
 If possible, prove a lower bound to this effect.
 Compare with the analogous situation for restricted smoothing splines, pointing out that the discretization imposed by the graph improves the situation slightly.
-	\item Study whether order-$m$ Laplacian smoothing estimate is optimal over $H^1(\mc{X})$, as long as $m > d/4$. 
-	\item Explain why your analysis of Laplacian smoothing fails at estimation over $H^s(\mc{X})$ for $s > 1$. If possible, prove a lower bound.
+	\item Study whether order-$m$ graph Laplacian smoothing estimate is optimal over $H^1(\mc{X})$, as long as $m > d/4$. 
+	\item Explain why your analysis of graph Laplacian smoothing fails at estimation over $H^s(\mc{X})$ for $s > 1$. If possible, prove a lower bound.
 \end{itemize}
@@
 
-Suppose now that in addition to the \emph{labeled} data $(x_1,y_1),\ldots,(x_n,y_n)$, we observe $m$ \emph{unlabeled} design points $x_{n + 1},\ldots,x_{n + m}$.
-Let $N = n + m$, and let $L_{N,\varepsilon}$ denote the graph Laplacian over labeled and unlabeled points,
-\begin{equation*}
+Suppose now that in addition to the \emph{labeled} data $(x_1,y_1),\ldots,(x_n,y_n)$, we observe $m$ \emph{unlabeled} design points $x_{n + 1},\ldots,x_{n + m}$ Let $N = n + m$, let $G_{N}$ denote the weighted graph over $x_1,\ldots,x_N$, and $L_{N,\varepsilon} \in \Reals^{N \times N}$ denote the graph Laplacian of $G_{N}$.
+Thus $L_{N,\varepsilon}$ satisfies,for any $u \in \Reals^N$,
+\begin{equation}
+\label{eqn:graph_laplacian_all}
 \bigl(L_{N,\varepsilon}u\bigr)_i := \frac{1}{N \varepsilon^{d + 2}}\sum_{j = 1}^{N} (u_i - u_j) \eta\biggl(\frac{\|x_i - x_j\|}{\varepsilon}\biggr).
-\end{equation*}
-The order-$m$ graph Laplacian smoothing estimator $\wh{f}$ defined in \textcolor{red}{(?)} can be naturally generalized to a vector in $\Reals^N$ according to the following formulation,
+\end{equation}
+The first-order graph Laplacian smoothing estimator $\wh{f}_n \in \Reals^n$ defined in \textcolor{red}{(?)} can be naturally generalized to an estimator $\wh{f}_N$ in $\Reals^N$ (that is, a solution over both labeled and unlabeled points), according to
 \begin{equation}
@@
 \end{equation}
-The estimator $\wh{f}$ has a closed-form solution, which can be simply written using the extension operator $E_{n \to N}: \Reals^{n} \to \Reals^N$ and its adjoint, the restriction operator $R_{N \to n}: \Reals^{N} \to \Reals^n$, which are defined by
+
+\paragraph{Two alternative forms.}
+\textcolor{red}{(TODO): Put in the Euler-Lagrange equation.}
+Letting the extension operator $E_{n \to N}: \Reals^{n} \to \Reals^N$ and its adjoint, the restriction operator $R_{N \to n}: \Reals^{N} \to \Reals^n$, be respectively defined as
 \begin{equation*}
-E_{n \to N}u := (u_1,\ldots,u_n,0,\ldots,0),\quad\textrm{and}\quad R_{N \to n}v := (v_1,\ldots,v_n).
+E_{n \to N}u := (u_1,\ldots,u_n,0,\ldots,0),\quad\textrm{and}\quad R_{N \to n}v := (v_1,\ldots,v_n),
 \end{equation*}
-Then the order-$m$ graph Laplacian smoothing estimator can be written as follows
+the first-order graph Laplacian smoothing estimator has the closed-form solution
 \begin{equation}
 \label{eqn:ssl_graph_laplacian_smoothing_1}
-\wh{f}_N = (E_{n \to N} R_{N \to n} + \frac{\rho n}{N} L_{N,\varepsilon})^{-1} E_{n \to N}Y.
+\wh{f}_N = \biggl(E_{n \to N} R_{N \to n} + \rho \cdot \frac{n}{N} L_{N,\varepsilon}\biggr)^{-1} E_{n \to N}Y.
 \end{equation}
 The form of~\eqref{eqn:ssl_graph_laplacian_smoothing_1} is a compact way to represent the solution $\wh{f}$, but it is rather opaque.
-In Proposition~\ref{prop:ssl_graph_laplacian_smoothing} we give a second form of $\wh{f}$, which is less compact but more evocative.
+In Proposition~\ref{prop:ssl_graph_laplacian_smoothing} we give a second form for $\wh{f}$, which is less compact but more evocative.
+To make sense of this alternative representation, we must introduce some new operators (matrices).
+Specifically, let $L_{m,\varepsilon} \in \Reals^{m \times m}$ be the graph Laplacian over only the unlabeled data, meaning for any vector $u \in \Reals^m$,
+\begin{equation}
+\label{eqn:graph_laplacian_unlabeled}
+\bigl(L_{m,\varepsilon}u\bigr)_i := \frac{1}{m \varepsilon^{d + 2}}\sum_{j = 1}^{m} (u_i - u_j) \eta\biggl(\frac{\|x_{n + i} - x_{n + j}\|}{\varepsilon}\biggr).
+\end{equation}
+Let $K_{m,n} \in \Reals^{m \times n}$ be defined entrywise as
+\begin{equation*}
+\bigl(K_{m,n}\bigr)_{ki} := \eta\biggl(\frac{\|x_i - x_{k + n}\|}{\varepsilon}\biggr).
+\end{equation*}
+Thus $(K_{m,n})_{ki}$ is simply the weight of the edge between the labeled point $x_i$ and the unlabeled point $x_{k +n}$ in the graph $G_N$.
+Let $D_{m,n} \in \Reals^{m \times m}$ be the diagonal matrix with entries $(D_{m,n})_{kk} := d_{n,\varepsilon}(x_{n + k}) := \sum_{j = 1}^{n} \bigl(K_{m,n}\bigr)_{n + k,j}$.
+Define $K_{n,m}$ and $D_{n,m}$ analogously.
+
+Finally, let $G_n^{(u)} := ([n], W_n^{(u)})$ be a weighted graph, with the $(i,j)$th entry of weighted edge matrix $W_n^{(u)}$ given by
+\begin{equation*}
+\bigl(W_{n}^{(u)}\bigr)_{ij} := \sum_{k,\ell = m + 1}^{m + n} \eta\biggl(\frac{\|x_i - x_k\|}{\varepsilon}\biggr) \cdot \eta\biggl(\frac{\|x_j - x_{\ell}\|}{\varepsilon}\biggr) \cdot \Bigl(m\varepsilon^{d + 2}L_{m,\varepsilon} + D_{m,n}\Bigr)_{k,\ell}^{-1}.
+\end{equation*}
+Let $L_{n,\varepsilon}^{(u)}$ be the corresponding Laplacian, meaning
+\begin{equation*}
+\bigl(L_{n,\varepsilon}^{(u)}u\bigr)_{i} := \frac{1}{m\varepsilon^{d + 2}} \sum_{j = 1}^{n} (u_i - u_j) W_{n,ij}^{(u)}.
+\end{equation*}
+The proof of Proposition~\ref{prop:ssl_graph_laplacian_smoothing} is contained in Section~\ref{subsec:pf_ssl_graph_laplacian_smoothing}. 
 \begin{proposition}
 	\label{prop:ssl_graph_laplacian_smoothing}
-	Let $\wh{f}_N$ be the solution to~\eqref{eqn:ssl_graph_laplacian_smoothing_0}. Writing $\wh{f}_N = (\wh{f}_n,\wh{f}_m)$ for $\wh{f}_n \in \Reals^m$ and $\wh{f}_n \in \Reals^m$, we have the following.
+	Let $\wh{f}_N$ be the solution to~\eqref{eqn:ssl_graph_laplacian_smoothing_0}.
+Writing $\wh{f}_N = (\wh{f}_N^{(\ell)},\wh{f}_N^{(u)})$ for $\wh{f}_N^{(\ell)} \in \Reals^n$ and $\wh{f}_N^{(u)} \in \Reals^m$, we have the following.
 	\begin{itemize}
-		\item 
+		\item The in-sample component $\wh{f}_N^{(\ell)}$ of the first-order Laplacian smoothing estimator $\wh{f}_N$ is given by
+		\begin{equation*}
+		\wh{f}_N^{(\ell)} := \biggl(I_n + \frac{\rho n}{N}\biggl[\frac{n}{N}L_{n,\varepsilon} + \frac{m}{N}L_{n,\varepsilon}^{(u)}\biggr]\biggr)^{-1} Y.
+		\end{equation*}
+		\item The out-of-sample component $\wh{f}_N^{(u)}$ of the first-order Laplacian smoothing estimator $\wh{f}_N$ is given by
+		\begin{equation*}
+		\wh{f}_N^{(u)} := \bigl(D_{m,n} + m\varepsilon^{d + 2}L_{m,\varepsilon}\bigr)^{-1} K_{m,n} \wh{f}_N^{(\ell)}.
+		\end{equation*}
 	\end{itemize}
 \end{proposition}
+Proposition~\ref{prop:ssl_graph_laplacian_smoothing} shows two things.
+First, it shows that at the labeled points $x_1,\ldots,x_n$, the estimator $\wh{f}_N$ is exactly an in-sample graph Laplacian smoothing estimator, but with the crucial distinction that the graph Laplacian is the (weighted) average of $L_{n,\varepsilon}$ and $L_{n,\varepsilon}^{(u)}$.
+The second thing Proposition~\ref{prop:ssl_graph_laplacian_smoothing} reveals is that at unlabeled points, the estimator $\wh{f}_N$ (or rather, its restriction to unlabeled points $\wh{f}_N^{(u)}$) satisfies the graph Poisson equation
+\begin{equation*}
+(D_{m,n} + m\varepsilon^{d + 2}L_{m,\varepsilon})\wh{f}_N^{(u)} = K_{m,n} \wh{f}_N^{(\ell)}.
+\end{equation*}
+In words, the out-of-sample component $\wh{f}_N^{(u)}$ is the solution to an inverse problem, in which the right hand side is a \emph{smoothening} of the in-sample component $\wh{f}_N^{(\ell)}$.
+This smoothening is a crucial aspect of the out-of-sample component.
+The role it plays is most clearly understood when there is exactly one unlabeled point.
+
+\subsection{One unlabeled point}
+Suppose $m = 1$.
+Then Proposition~\ref{prop:ssl_graph_laplacian_smoothing} simplifies.
+In particular, the out-of-sample component of $\wh{f}_N$ can be written in terms of the Nadaraya-Watson kernel smoother $T_{n,\varepsilon}$ defined by $T_{n,\varepsilon}u(x) := 1/d_{n,\varepsilon}(x) K_{(x)} u :=  1/d_{n,\varepsilon}(x) \cdot \sum_{i = 1}^{n} u_i \eta(\|x_i - x\|/\varepsilon)$.
+Here we have written $K_{(x)}$ for the column vector of kernel evaluations at labeled points, $K_{(x)} := (\eta(\|x_1 - x\|/\varepsilon),\ldots,\eta(\|x_n - x\|/\varepsilon))$.
+\begin{corollary}
+	\label{cor:ssl_graph_laplacian_smoothing_one_unlabeled}
+	Let $\wh{f}_N$ be the solution to~\eqref{eqn:ssl_graph_laplacian_smoothing_0}.
+Writing $\wh{f}_N = (\wh{f}_N^{(\ell)},\wh{a})$ for $\wh{f}_N^{(\ell)} \in \Reals^n$ and $\wh{a} \in \Reals$, we have the following.
+	\begin{itemize}
+		\item The in-sample component $\wh{f}_N^{(\ell)}$ of the first-order Laplacian smoothing estimator $\wh{f}_N$ is given by
+		\begin{equation}
+		\label{eqn:ssl_graph_laplacian_smoothing_one_unlabeled_1}
+		\wh{f}_N^{(\ell)} := \biggl(I_n + \frac{\rho n}{n + 1}\biggl[\frac{n}{n + 1}L_{n,\varepsilon} + \frac{1}{n + 1}L_{n,\varepsilon}^{(1)}\biggr]\biggr)^{-1} Y.
+		\end{equation}
+		\item The out-of-sample component $\wh{a}$ of the first-order Laplacian smoothing estimator $\wh{f}_N$ is given by
+		\begin{equation}
+		\label{eqn:ssl_graph_laplacian_smoothing_one_unlabeled_2}
+		\wh{a} = \bigl(T_{n,\varepsilon}\wh{f}_N^{(\ell)}\bigr)(x_{n + 1}) = \frac{1}{d_{n,\varepsilon}(x_{n + 1})} K_{(x_{n + 1})}\wh{f}_N^{(\ell)}.
+		\end{equation}
+	\end{itemize}
+\end{corollary}
+From~\eqref{eqn:ssl_graph_laplacian_smoothing_one_unlabeled_1} we see that, roughly speaking, as long as $L_{n,\varepsilon}^{(u)}$ has a reasonable spectrum, the in-sample component $\wh{f}_N^{(\ell)}$ should be quite close to the Laplacian smoothing estimator $\wh{f}_n$ formed only using the in-sample points.
+From~\eqref{eqn:ssl_graph_laplacian_smoothing_one_unlabeled_2} we see that the out-of-sample component $\wh{a}$ is constructed by passing a Nadaraya-Watson kernel smoother (with kernel function equal to $\eta$ and bandwidth equal to $\varepsilon$) over $\wh{f}_N^{(\ell)}$.
+
+\section{Proofs}
+
+\subsection{Proof of Proposition~\ref{prop:ssl_graph_laplacian_smoothing}}
+\label{subsec:pf_ssl_graph_laplacian_smoothing}
+\textcolor{red}{(TODO)}
+
+\subsection{Proof of Theorem~{(?)}}
+\label{subsec:pf_ssl_graph_laplacian_smoothing_2}
+
+\paragraph{Strategy.}
+The decomposition in Proposition~\ref{prop:ssl_graph_laplacian_smoothing} reveals something of the structure of $\wh{f}_N$.
+It also suggests a route to analyzing the out-of-sample error $(\wh{a} - f_0(x_{n + 1}))^2$.
+Concretely, consider the following expansion
+\begin{align*}
+\wh{a} - f_0(x_{n + 1}) & = \wh{a} - \frac{1}{d_{n,\varepsilon}(x_{n + 1})}K_{(x_{n + 1})} \Ebb_n\wh{f}_N^{(\ell)} + \\
+& \quad \frac{1}{d_{n,\varepsilon}(x_{n + 1})}K_{(x_{n + 1})} \Ebb_n\wh{f}_N^{(\ell)} - \frac{1}{d_{n,\varepsilon}(x_{n + 1})}K_{1,n} f_0 + \\
+& \quad \frac{1}{d_{n,\varepsilon}(x_{n + 1})}K_{(x_{n + 1})} f_0 - f_0(x_{n + 1}) \\
+\Longrightarrow \Ebb_n\bigl[|\wh{a} - f_0(x_{n + 1})|^2\bigr]& \leq \underbrace{\|T_{n,\varepsilon}(\wh{f}_N^{(\ell)} - \mathbb{E}_{n}\wh{f}_N^{(\ell)})\|_P^2}_{\textrm{variance}} + \underbrace{\|T_{n,\varepsilon}(\mathbb{E}_{n}\wh{f}_N^{(\ell)} - f_0)\|_P^2}_{\textrm{bias}} + \underbrace{\|T_{n,\varepsilon}f_0 - f_0\|_P^2}_{\textrm{noiseless kernel smoothing}}
+\end{align*}
+Here $\Ebb_n$ denotes the expectation conditional on the labeled design points $x_1,\ldots,x_n$. Now we proceed to bound each of these terms, with high probability.
+
+\paragraph{Term 1: Variance.}
+Let $\wt{L}_{n,\varepsilon} = n/(n + 1) L_{n,\varepsilon} + 1/(n + 1) L_{n,\varepsilon}^{(u)}$.
+Denoting the smoother matrix $S_n := [I_n + \rho n/(n + 1)\wt{L}_{n,\varepsilon}]^{-1}$, we can write the variance term as
+\begin{equation*}
+\|T_{n,\varepsilon}(\wh{f}_N^{(\ell)} - \mathbb{E}_{n}\wh{f}_N^{(\ell)})\|_P^2 =  \int_{\Omega} \Bigl(T_{n,\varepsilon}(S_nw)(x)\Bigr)^2 \,dP(x).
+\end{equation*}
+
+Note that for any $i,j$ in $1,\ldots,n$, we have
+\begin{equation*}
+\Ebb_n\bigl[(S_nw)_i \cdot (S_nw)_j\bigr] = (S_n^2)_{ij}.
+\end{equation*}
+Therefore, at a given $x \in \Omega$ it follows that
+\begin{equation*}
+\Ebb_n\Bigl[\bigl(T_{n,\varepsilon}S_nw(x)\bigr)^2\Bigr] = \frac{1}{d_{n,\varepsilon}(x)^2} K_{(x)}^{\top} S_n^2 K_{(x)} \leq \frac{1}{d_{n,\varepsilon}(x)^2} \|S_n\|_{\textrm{op}}^2 \cdot \|K_{(x)}\|_2^2 \leq \frac{1}{d_{n,\varepsilon}(x)^2} \cdot \|K_{(x)}\|_2^2.
+\end{equation*}
+Now by Lemma~\ref{lem:max_degree_deviation}, we have that for both $K(x) = \eta(x)$ and $K = \bigl(\eta(x)\bigr)^2$, with probability at least $1 - C_2\exp\bigl(-c_2 \delta^2 \cdot n \varepsilon^d\bigr)$,
+\begin{equation}
+\label{pf:out_of_sample_laplacian_regularization_1}
+\sup_{x \in \Omega} \biggl|\frac{d_{n,\varepsilon}(x; K)}{n} - d_{P,\varepsilon}(x; K)\biggr| \leq \delta \cdot \inf_{x \in \Omega} d_{P,\varepsilon}(x; K),
+\end{equation}
+and thus
+\begin{equation*}
+\sup_{x \in \Omega}\frac{1}{d_{n,\varepsilon}(x)^2} \cdot \|K_{(x)}\|_2^2 \leq \biggl[\frac{(1 + \delta)}{(1 - \delta)}\biggr]^2 \sup_{x \in \Omega} \frac{1}{nd_{P,\varepsilon}(x)^2} \int \bigl[\eta(\|x'-x\|/\varepsilon)\bigr]^2 \,dP(x') \leq \biggl[\frac{(1 + \delta)}{(1 - \delta)}\biggr]^2 \frac{\|\eta\|_{\infty}}{n p_{\min} \varepsilon^d}.
+\end{equation*}
+Integrating with respect to $P$, and exchanging expectation with integration thus gives that with probability at least $1 - C_2\exp\bigl(-c_2 \delta^2 \cdot n \varepsilon^d\bigr)$,
+\begin{equation*}
+\|T_{n,\varepsilon}(\wh{f}_N^{(\ell)} - \mathbb{E}_{n}\wh{f}_N^{(\ell)})\|_P^2 \leq \biggl[\frac{(1 + \delta)}{(1 - \delta)}\biggr]^2 \frac{\|\eta\|_{\infty} p_{\max}}{n p_{\min} \varepsilon^d}.
+\end{equation*}
+
+\paragraph{Term 2: Bias.}
+\textcolor{red}{(TODO)}.
+
+\paragraph{Term 3: Noiseless kernel smoothing.}
+This is a standard analysis. \textcolor{red}{From our paper on Laplacian eigenmaps (equation (107) and Lemma~19)}, we have that
+\begin{equation*}
+\|T_{n,\varepsilon}f_0 - f_0\|_P^2 \leq \frac{C}{\delta} \cdot \biggl(\frac{\varepsilon^2}{n\varepsilon^d} + \varepsilon^2\biggr) |f|_{H^1(\Omega)}^2 \leq \frac{C}{\delta} \cdot \varepsilon^2 |f|_{H^1(\Omega)}^2
+\end{equation*}
+with probability at least $1 - C\exp(-cn\varepsilon^d)$.
+
+\section{Technical Results}
+
+\subsection{Concentration of Empirical Functionals}
+
+\paragraph{Degree.}
+Let $d_{n,h}(x;K) := \sum_{i = 1}^{n} K(\|x_i - x\|/h)$ be the empirical degree functional of a kernel function $K$.
+We begin by recalling some results the concentration of the degree functional $d_{n,h}(x)$ around its mean.
+This concentration statement will be meaningful whenever $K$ is a smooth kernel and $h$ belongs to some range.
+\begin{enumerate}[label=(K\arabic*)]
+	\setcounter{enumi}{0}
+	\item
+	\label{asmp:kernel}
+	The kernel function $K$ is supported on a subset of $[0,1]$. Additionally, $K$ is Lipschitz continuous on $[0,1]$, and is normalized so that
+	\begin{equation*}
+	\int_{-\infty}^{\infty} K(|z|) \,dz = 1.
+	\end{equation*}
+\end{enumerate}
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{0}
+	\item
+	\label{asmp:bandwidth}
+	For $\delta > 0$, and for constants $c_0$ and $C_0$, the bandwidth parameter $h$ satisfies
+	\begin{equation*}
+	\frac{1}{\delta^2}C_0\biggl(\frac{\log(1/h)}{n}\biggr)^{1/d} \leq h \leq c_0.
+	\end{equation*}
+\end{enumerate}
+The constants $C_0$ and $c_0$ will be specified in the proof of the following Lemma.
+\begin{lemma}
+	\label{lem:max_degree_deviation}
+	Fix $\delta \in (0,1)$ \textcolor{red}{sufficiently small}.
+Suppose $\Omega$ and $P$ are \textcolor{red}{regular}, and that the kernel $K$ satisfies~\ref{asmp:kernel} and the bandwidth $h$ satisfies~\ref{asmp:bandwidth}.
+It follows that
+	\begin{equation*}
+	\Pbb\Biggl( \sup_{x \in \Rd} \biggl|\frac{d_{n,h}(x)}{n} - d_{P,h}(x)\biggr| \geq \delta \cdot \min_{x \in \Omega}\bigl\{d_{P,h}(x)\bigr\} \Biggr) \leq C_2\exp\bigl(-c_2 \delta^2 \cdot n h^d\bigr)
+	\end{equation*}
+\end{lemma}
+\begin{proof}
+	\textcolor{red}{(TODO)} Refer to Lemma~23 of Laplacian Eigenmaps paper. 
+\end{proof}
 

### 2021-05-21 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Added proof of kernel-smoothing of in-sample estimator.
Cosmetic changes to Laplacian Eigenmaps manuscript.

--- 2021-04-28 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
@@
 	\begin{equation*}
-	C_0\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq h \leq c_0.
+	C_0\biggl(\frac{\log(1/h)}{n}\biggr)^{1/d} \leq h \leq c_0.
 	\end{equation*}
@@
 
-\textcolor{red}{(TODO): I list some potential options below.}
-\begin{itemize}
-	\item \textbf{Laplacian eigenmaps approximates spectral projection.} $1$d unit cube, uniform density, ``interesting'' regression function.
-Compare Laplacian eigenmaps with spectral projection estimator, including their MSE curves, their fits at several different values of $K$, and their risk as a function of $n$.
-	\item \textbf{Laplacian eigenmaps beats nonparametric least squares.} $1$d unit cube, very non-uniform density, same regression function.
-Compare risk of Laplacian eigenmaps to nonparametric least squares (with eigenfunctions of unweighted Laplace operator as features) as a function of $n$.
-Show that eigenmaps outperforms least squares.
-	\item \textbf{Investigate our theoretical requirements for hyperparameters.} $1$d unit cube, uniform density, ``smooth'' regression function.
-Investigate the MSE heat map of Laplacian eigenmaps, as a function of $K$ and $\varepsilon$.
-In particular, see if the best combination of these hyperparameters involves a relatively large choice of bandwidth $\varepsilon$.
-Possibly separately plot condtional-on-$X$ bias and variance, as a function of $K$, for a few different choices of $\varepsilon$.
-	\item \textbf{Manifold setting.} Swiss roll, density \textcolor{red}{(?)}, regression function \textcolor{red}{(?)}.
-Evaluate Laplacian eigenmaps in-sample error, and Laplacian-eigenmaps + kernel smoothing out-of-sample error.
-	\item \textbf{Laplacian eigenmaps beats Laplacian smoothing.} High-dimensional unit cube, uniform density, ``interesting'' regression function.
-Compare risk of Laplacian eigenmaps to risk of Laplacian smoothing as a function of $n$, show that the former outperforms the latter.
-\end{itemize}
+\textcolor{red}{(TODO): Include more details.}
+In this section we empirically demonstrate that Laplacian Eigenmaps is a very good alternative to spectral projection, even when $n$ is only moderately large.
+In order to compare the two methods, in our experiments we stick to simple settings where we can compute eigenfunctions of $\Delta_P$, and thus the spectral projection estimator.
+In general, this is not so easily done: hence the appeal of Laplacian Eigenmaps.
+
+In our first experiment, we compare the mean-squared error of Laplacian eigenmaps to that of its classical counterpart.
+We vary the sample size from $n = 200$ to $n = 2000$; sample $n$ design points $X_1,\ldots,X_n$ from the uniform distribution on the cube $[-1,1]^d$; and sample responses $Y_i$ according to~\eqref{eqn:model} with regression function $f_0 = M/\sqrt{n} \cdot \sum_{k = 1}^{n/20} k^{-s/d} \psi_k$ (Note that $f_0$ changes with $n$, but always belongs to the Sobolev ball~\textcolor{red}{(?)}.) In Figure~\ref{fig:fig1} we show the mean-squared error of Laplacian eigenmaps and spectral projection as a function of $n$, for different dimensions $d$ and order of smoothness $s$.
+We see that both estimators have mean-squared error converging to zero at least as fast as the optimal rate.
+(Here the regression function $f_0$ is a little bit smoother than the worst-case Sobolev function, which explains why the minimax rate is not always saturated).
+Surprisingly, we also find that Laplacian Eigenmaps slightly but consistently outperforms its continuum counterpart.
+\begin{figure*}[tb]
+	\includegraphics[width=.245\textwidth]{figures/mse_by_sample_size_2d.pdf}
+	\includegraphics[width=.245\textwidth]{figures/mse_by_sample_size_3d.pdf} 
+	\includegraphics[width=.245\textwidth]{figures/mse_by_sample_size_2d_2s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/mse_by_sample_size_3d_2s.pdf} 
+	\caption{Mean squared error of Laplacian eigenmaps (\texttt{LE}) and a spectral projection estimator (\texttt{SP}) as a function of sample size $n$.
+Each plot is on the log-log scale, and the results are averaged over 50 repetitions.
+Both estimators are tuned for optimal average mean squared error.
+The black line shows the minimax rate (in slope only; the intercept is chosen to match the observed error).}
+	\label{fig:fig1}
+\end{figure*}
+
+In Figure~\ref{fig:fig2} we compare tests using Laplacian eigenmaps and spectral projection. \textcolor{red}{(TODO): Need to ask Siva for help.}
+\begin{figure*}[tb]
+	\caption{Worst-case risk Laplacian eigenmaps (\texttt{LE}) and spectral projection (\texttt{SP}) tests, as a function of sample size $n$.}
+	\label{fig:fig2}
+\end{figure*}
+
+These experiments demonstrate that in terms of statistical error, Laplacian eigenmaps methods are reasonable replacements for spectral projection methods.
+Laplacian eigenmaps depends on two tuning parameters, and in our final experiment we investigate the importance of both, focusing now on estimation.
+In Figure~\ref{fig:fig3}, we see how the mean-squared error of Laplacian eigenmaps changes as each tuning parameter is varied.
+As suggested by our theory, properly choosing the number of eigenvectors $K$ is crucial: the mse curves have a sharply defined minimum.
+On the other hand, as a function of the graph radius parameter $\varepsilon$ the mse curves are much closer to flat.
+This is particularly true in the first-order ($s = 1$) case, and still true albeit to a lesser extent when $s = 2$.
+These findings reinforce two points we've made previously: first that Laplacian Eigenmaps should be viewed as a type of spectral projection method, in which the more important parameter to carefully tune is the number of eigenvectors $K$; second that the more regularity displayed by the regression function, the more important tuning the graph radius $\varepsilon$ becomes.
+
+\begin{figure*}[tb]
+	\includegraphics[width=.245\textwidth]{figures/eigenmaps_parameters/mse_by_number_of_eigenvectors.pdf}
+	\includegraphics[width=.245\textwidth]{figures/eigenmaps_parameters/mse_by_radius.pdf} 
+	\includegraphics[width=.245\textwidth]{figures/eigenmaps_parameters_2s/mse_by_number_of_eigenvectors.pdf}
+	\includegraphics[width=.245\textwidth]{figures/eigenmaps_parameters_2s/mse_by_radius.pdf} 
+	\caption{Mean squared error of Laplacian Eigenmaps as a function of hyperparameters.
+For all experiments, sample size $n = 2000$, and the results are averaged over $50$ repetitions.
+Colors show~\textcolor{red}{(?)}}
+	\label{fig:fig3}
+\end{figure*}
 

### 2021-05-21 code/configs/laplacian_eigenmaps/eigenmaps_approximates_spectral_projection.R
Updated code base for running Laplacian regression simulations.

@@ -1,20 +1,0 @@
-#------------------------------------------------#
-# Configs for Laplacian Eigenmaps Figure 1.
-#------------------------------------------------#
-
-### General configs.
-d <- 1
-# ns <- 2000
-ns <- round(seq(400,2000,length.out = 10))
-iters <- 50
-
-### Configs for sampling. ###
-sample_X <- make_sample_uniform(d)
-make_f0 <- make_spectral_sobolev
-
-
-### Methods. ###
-methods <- list(laplacian_eigenmaps = make_laplacian_eigenmaps,
-                spectral_projection = make_spectral_projection)
-initialize_thetas <- list(initialize_laplacian_eigenmaps_thetas,
-                          initialize_spectral_projection_thetas)

### 2021-05-21 code/estimation_mse_pipeline.R

@@ -6,12 +6,15 @@ library(RSpectra)
 source("sample.R")
 source("graph.R")
 source("estimators.R")
+source("manifold.R")
+source("misc.R")
 
 # Please change this line to whichever config you wish to run.
-source("configs/laplacian_eigenmaps/eigenmaps_approximates_spectral_projection.R")
+source("configs/laplacian_eigenmaps/mse/sobolev_2d_2s.R")
 
-# Please change this to dictate how verbose the pipeline should be 
+# Options for running the pipeline
 verbose <- F
+save_fits <- F
 
 #----------------------------------------------------#
 # This is the pipeline for running regression estimation experiments.
@@ -24,14 +27,17 @@ Xs <- vector(mode = "list", length = length(ns))
 f0s <- vector(mode = "list", length = length(ns))
 Ys <- vector(mode = "list", length = length(ns))
 thetas <- vector(mode = "list", length = length(ns))
-fits <- vector(mode = "list", length = length(ns))
+if(save_fits) fits <- vector(mode = "list", length = length(ns))
 
 for(ii in 1:length(ns))
 {
   n <- ns[ii]
   f0 <- make_f0(d,n)
   mse[[ii]] <- vector(mode = "list", length = length(methods))
-  fits[[ii]] <- vector(mode = "list",length = length(methods))
+  if(save_fits)
+  {
+    fits[[ii]] <- vector(mode = "list",length = length(methods))
+  }
 
   # Initialize tuning parameters (thetas).
   # Note: I have granted methods the right to access the distribution P from which X is sampled
@@ -41,7 +47,7 @@ for(ii in 1:length(ns))
   {
     thetas[[ii]][[jj]] <- initialize_thetas[[jj]](sample_X, n)
     mse[[ii]][[jj]] <- matrix(nrow = nrow(thetas[[ii]][[jj]]), ncol = iters)
-    fits[[ii]][[jj]] <- array(dim = c(nrow(thetas[[ii]][[jj]]),n,iters))
+    if(save_fits) fits[[ii]][[jj]] <- array(dim = c(nrow(thetas[[ii]][[jj]]),n,iters))
   }
   for(iter in 1:iters)
   {
@@ -91,7 +97,7 @@ for(ii in 1:length(ns))
     }
     
     ### Save fits.
-    fits[[ii]][[jj]][,,iter] <- fits_by_method[[jj]]
[…]

### 2021-05-21 code/estimators.R

@@ -22,8 +22,6 @@ make_laplacian_eigenmaps <- function(theta)
     a_k <- t(V_K) %*% Y         # Empirical Fourier coefficients
     f_hat <- V_K %*% a_k        # Spectral projection
     
-    # browser()
-    
     return(f_hat)
   }
 }
@@ -92,13 +90,28 @@ make_spectral_projection <- function(theta)
   spectral_projection <- function(Y,X)
   {
     d <- ncol(X)
-    stopifnot(d == 1)
-    
     n <- length(Y)
-    psi_K <- sapply(0:(K - 1),function(k){sqrt(2) * cos(k*pi*(X[,1] + 1)/2)}) # Fourier basis
-    a_k <- 1/n * t(psi_K) %*% Y   # Empirical Fourier coefficients
+  
+    if(exists("precomputed_data")){
+      stopifnot(length(precomputed_data) == 1)
+      psi_K <- precomputed_data[[1]][,1:K,drop = FALSE]
+    } else{
+      trig_basis <- get_trigonometric_basis(d,K) # Fourier basis
+      psi_K <- if(K == 1) apply(X,1,trig_basis) else apply(X,1,trig_basis) %>% t()
+    }
+    a_k <- 1/n * (t(psi_K) %*% Y)   # Empirical Fourier coefficients
     f_hat <- psi_K %*% a_k        # Spectral projection
   }
+} 
+attr(make_spectral_projection,"precompute_data") <- function(Y,X, theta_df){
+  precomputed_data <- vector(mode = "list",length = 1)
+  
+  K <- max(theta_df$K)
+  trig_basis <- get_trigonometric_basis(d,K) # Fourier basis
+  psi_K <- if(K == 1) apply(X,1,trig_basis) else apply(X,1,trig_basis) %>% t()
+  
+  precomputed_data[[1]] <- psi_K
+  return(precomputed_data)
 }
 
 make_least_squares <- function(theta)
@@ -107,12 +120,28 @@ make_least_squares <- function(theta)
   least_squares <- function(Y,X)
   {
     d <- ncol(X)
-    stopifnot(d == 1)
-    
     n <- length(Y)
-    psi_K <- sapply(0:(K - 1),function(k){sqrt(2) * cos(k*pi*(X[,1] + 1)/2)}) # Fourier basis
-    f_hat <- lm.fit(x = psi_K,y = Y)$fitted.values              # Least squares
+    
[…]

### 2021-05-21 code/graph.R

@@ -1,10 +1,9 @@
 neighborhood_graph <- function(X, r)
 {
   # unpack necessary parameters
-  d <- ncol(X)
   N <- nrow(X)
   
-  k_max <- min(round(N*pi^{d/2}/(gamma(d/2 + 1))*r^d*2),N - 1)
+  k_max <- round(min(4*N^(.5),N - 1))
   A <- knn_to_neighborhood_graph(X,r,k_max)
   return(A)
 }

### 2021-05-21 code/plot_demos/eigenmaps_figure_1.R

@@ -1,155 +1,0 @@
-library(dplyr)
-library(ggplot2)
-library(gridExtra)
-library(viridis)
-source("plot_methods.R")
-
-# User entered information.
-data_directory <- "data/laplacian_eigenmaps/20210511125310"
-plot_directory <- "plots/laplacian_eigenmaps/eigenmaps_approximates_spectral_projection" # Please change this to whichever directory you prefer.
-if(!exists(plot_directory)){dir.create(plot_directory)}
-
-# Load data
-load(file.path(data_directory,"configs.R"))
-d <- configs$d; ns <- configs$ns; methods <- configs$methods
-load(file.path(data_directory,"best_fits_by_method.R"))
-load(file.path(data_directory,"thetas.R"))
-load(file.path(data_directory,"mse.R"))
-load(file.path(data_directory,"Xs.R"))
-load(file.path(data_directory,"f0s.R"))
-load(file.path(data_directory,"Ys.R"))
-
-# Plotting parameters for all plots
-
-## Plot 1: Regression fxn and fitted values.
-plot1_indx <- 5
-plot1_K <- 10
-rs <- unique(thetas[[plot1_indx]][[1]]$r)
-plot1_r <- rs[which.min(abs(rs - .02))]    # arbitrary choice
-
-# Regression fxn
-pdf(file.path(plot_directory,"f0.pdf"))
-par(mar = c(5.1,6,4.1,4.1)) # Prevent the left hand side from getting cut off.
-plot_fxn(Xs[[plot1_indx]],Ys[[plot1_indx]],f = f0s[[plot1_indx]])
-dev.off()
-
-# Eigenmaps fitted values
-pdf(file.path(plot_directory,"eigenmaps.pdf"))
-par(mar = c(5.1,6,4.1,4.1)) # Prevent the left hand side from getting cut off.
-f <- subset(fits[[plot1_indx]][[1]], 
-            thetas[[plot1_indx]][[1]]$r == plot1_r & 
-            thetas[[plot1_indx]][[1]]$K == plot1_K)[1,]
-            
-plot_fxn(Xs[[plot1_indx]],Ys[[plot1_indx]],f,title = "Eigenmaps, K = 10",col = "red")
-dev.off()
-
-# Eigenmaps fitted values
[…]

### 2021-05-21 code/plot_methods.R

@@ -12,6 +12,7 @@ plot_fxn <- function(x,y,f,title = NULL,col = NULL){
 
 # Plot of mse---for best choice of tuning parameter---by sample size
 plot_best_mse <- function(methods,mse){
+  stopifnot(length(methods) <= 3)
   cols <- c("red","blue","green")[1:length(methods)]     
   
   plot_dfs_best_mse <- vector(mode = "list", length = length(methods))
@@ -30,7 +31,6 @@ plot_best_mse <- function(methods,mse){
       sd_best_mse[ii] <- apply(mse_ii_jj,1,sd)[which.min(rowMeans(mse_ii_jj))]/sqrt(ncol(mse_ii_jj))
       minimax_mse[ii] <- ns[ii]^{-2/(2 + d)}
     }
-    
     # Rescale minimax mse to match intercept with best_mse
     minimax_mse <- minimax_mse * (best_mse[1]/minimax_mse[1])
     plot_dfs_best_mse[[jj]] <- data.frame(x = ns, y = best_mse,sd = sd_best_mse) 
@@ -51,7 +51,12 @@ plot_best_mse <- function(methods,mse){
                                              " [Slope = ", fitted_slopes[jj],"].")
   }
   
-  title <- paste0("d = ", d,". Minimax slope = ", -2,"/", d+2, ".")
+  # browser()
+  # # ALDEN REMOVE
+  # best_K <- find_best_K(mse,thetas)
+  # #
+  
+  title <- paste0("d = ", d,", s = ",s,".", "Minimax slope = ", -2*s,"/", d+2*s, ".")
   plot_df_best_mse <- bind_rows(plot_dfs_best_mse, .id = "method") 
   legend_text <- unique(plot_df_best_mse$method)
   
@@ -62,7 +67,8 @@ plot_best_mse <- function(methods,mse){
   
   # shadow plot
   plot(x = ns, xlim = xlims, ylim = ylims,
-       log = "xy", xlab = "Sample size", ylab = "Mean squared error", main = title, 
+       log = "xy", xlab = "Sample size", ylab = "Mean squared error", 
+       main = title,
        cex.main = 2.5, cex.lab = 2.5, cex.axis = 2)
   
   # add points and lines
@@ -85,6 +91,6 @@ plot_best_mse <- function(methods,mse){
   # 
   lines(x = ns, y = minimax_mse)
   grid(equilogs = F, lwd = 2)
[…]

### 2021-05-21 code/sample.R

@@ -34,6 +34,15 @@ make_sample_uniform <- function(d)
   }
 }
 
+make_sample_manifold <- function(d,D)
+{
+  stopifnot(d <= D)
+  sample_X <- function(n){
+    Z <- matrix(runif(n*d,-1,1),ncol = d)
+    X <- t(apply(Z,1,FUN = function(z){chart(z,D)}))
+  }
+}
+
 #-----------------------------------------------#
 # Regression functions
 #-----------------------------------------------#
@@ -46,19 +55,31 @@ make_doppler <- function(d,n){
   }
 }
 
-make_spectral_sobolev <- function(d,n){
+make_spectral_sobolev <- function(d,n,s = 1,M = 2^s,beta = NULL){
   #------------------------------#
   # Fixes a vector beta that belongs to a Sobolev ellipsoid,
   # then builds the regression function
   # 
   # f0(x) = sum_{k} beta_k psi_k
   #
-  # where psi_k(x) = cos(k*pi*x)
-  stopifnot(d == 1) # Only do 1d for the moment
-  c <- sqrt(sum((1:3)^2))
-  beta <- c(rep(1/c,3),1/(sqrt(17)*(4:20)))
+  # where psi_k(x) is the kth element in the trigonometric basis.
+  #------------------------------#
+  # c <- sqrt(sum((1:3)^(2/d)))
+  # beta <- c(0,rep(1/c,3),1/(sqrt(17)*sqrt((4:20)^(2/d))))
+  if(is.null(beta)){
+    n_coefs <- n/10
+    beta <- M/sqrt(n_coefs) * c(1/sqrt( (1:n_coefs)^(2*s/d) ) ) # Sobolev norm = 2^s.
+  }
+  trig_basis <- get_trigonometric_basis(d,length(beta))
   f0 <- function(x){
-    sum(sapply(1:length(beta), function(idx){beta[idx] * sqrt(2) * cos(idx/2*pi*(x + 1))}))
+    t(beta) %*% trig_basis(x)
+  }
+}
+
+make_spectral_sobolev_manifold <- function(d,n,beta = NULL){
+  g0 <- make_spectral_sobolev(d,n,beta)
+  f0 <- function(x){
+    g0(inverse_chart(x,d))
   }
 }
 

### 2021-05-21 code/configs/laplacian_eigenmaps/eigenmaps_tuning_2s.R
Added many configs for creating figures for Laplacian Eigenmaps experiments section.

@@ -1,0 +1,21 @@
+#------------------------------------------------#
+# Configs for Laplacian Eigenmaps Figure 1.
+#------------------------------------------------#
+
+### General configs.
+d <- 2
+ns <- 1000
+# ns <- round(seq(400,2000,length.out = 10))
+iters <- 50
+
+### Configs for sampling. ###
+sample_X <- make_sample_uniform(d)
+s <- 2
+make_f0 <- function(d,n){make_spectral_sobolev(d,n,s)}
+
+
+### Methods. ###
+methods <- list(laplacian_eigenmaps = make_laplacian_eigenmaps,
+                spectral_projection = make_spectral_projection)
+initialize_thetas <- list(initialize_laplacian_eigenmaps_thetas,
+                          initialize_spectral_projection_thetas)

### 2021-05-21 code/configs/laplacian_eigenmaps/mse/sobolev_2d_1s.R

@@ -1,0 +1,20 @@
+#------------------------------------------------#
+# Configs for Laplacian Eigenmaps Figure 1.
+#------------------------------------------------#
+
+### General configs.
+d <- 2
+s <- 1
+ns <- round(seq(400,2000,length.out = 10))
+iters <- 50
+
+### Configs for sampling. ###
+sample_X <- make_sample_uniform(d)
+make_f0 <- function(d,n){make_spectral_sobolev(d,n,s)}
+
+
+### Methods. ###
+methods <- list(laplacian_eigenmaps = make_laplacian_eigenmaps,
+                spectral_projection = make_spectral_projection)
+initialize_thetas <- list(initialize_laplacian_eigenmaps_thetas,
+                          initialize_spectral_projection_thetas)

### 2021-05-21 code/configs/laplacian_eigenmaps/mse/sobolev_2d_2s.R

@@ -1,0 +1,24 @@
+#------------------------------------------------#
+# Configs for Laplacian Eigenmaps Figure 1.
+#------------------------------------------------#
+
+### General configs.
+d <- 2
+s <- 2
+M <- 16
+# ns <- 2000
+ns <- round(seq(400,2000,length.out = 10))
+iters <- 50
+
+### Configs for sampling. ###
+sample_X <- make_sample_uniform(d)
+make_f0 <- function(d,n){make_spectral_sobolev(d,max(ns),s,M)}
+
+
+### Methods. ###
+methods <- list(laplacian_eigenmaps = make_laplacian_eigenmaps,
+                spectral_projection = make_spectral_projection,
+                least_squares = make_least_squares)
+initialize_thetas <- list(initialize_laplacian_eigenmaps_thetas,
+                          initialize_spectral_projection_thetas,
+                          initialize_spectral_projection_thetas)

### 2021-05-21 code/configs/laplacian_eigenmaps/mse/sobolev_3d_1s.R

@@ -1,0 +1,20 @@
+#------------------------------------------------#
+# Configs for Laplacian Eigenmaps Figure 1.
+#------------------------------------------------#
+
+### General configs.
+d <- 3
+s <- 1
+ns <- round(seq(400,2000,length.out = 10))
+iters <- 50
+
+### Configs for sampling. ###
+sample_X <- make_sample_uniform(d)
+make_f0 <- function(d,n){make_spectral_sobolev(d,n,s)}
+
+
+### Methods. ###
+methods <- list(laplacian_eigenmaps = make_laplacian_eigenmaps,
+                spectral_projection = make_spectral_projection)
+initialize_thetas <- list(initialize_laplacian_eigenmaps_thetas,
+                          initialize_spectral_projection_thetas)

### 2021-05-21 code/configs/laplacian_eigenmaps/mse/sobolev_3d_2s.R

@@ -1,0 +1,22 @@
+#------------------------------------------------#
+# Configs for Laplacian Eigenmaps Figure 1.
+#------------------------------------------------#
+
+### General configs.
+d <- 3
+s <- 2
+ns <- round(seq(400,2000,length.out = 10))
+iters <- 50
+
+### Configs for sampling. ###
+sample_X <- make_sample_uniform(d)
+make_f0 <- function(d,n){make_spectral_sobolev(d,n,s)}
+
+
+### Methods. ###
+methods <- list(laplacian_eigenmaps = make_laplacian_eigenmaps,
+                spectral_projection = make_spectral_projection,
+                least_squares = make_least_squares)
+initialize_thetas <- list(initialize_laplacian_eigenmaps_thetas,
+                          initialize_spectral_projection_thetas,
+                          initialize_spectral_projection_thetas)

### 2021-05-21 code/configs/laplacian_eigenmaps/mse/sobolev_4d_1s.R

@@ -1,0 +1,21 @@
+#------------------------------------------------#
+# Configs for Laplacian Eigenmaps Figure 1.
+#------------------------------------------------#
+
+### General configs.
+d <- 4
+s <- 1
+ns <- round(seq(400,2000,length.out = 10))
+iters <- 200
+
+### Configs for sampling. ###
+sample_X <- make_sample_uniform(d)
+make_f0 <- function(d,n){make_spectral_sobolev(d,n,s)}
+
+
+### Methods. ###
+methods <- list(laplacian_eigenmaps = make_laplacian_eigenmaps,
+                spectral_projection = make_spectral_projection,
+                least_squares = make_least_squares)
+initialize_thetas <- list(initialize_laplacian_eigenmaps_thetas,
+                          initialize_spectral_projection_thetas)

### 2021-05-21 code/configs/laplacian_eigenmaps/mse/sobolev_manifold_3d.R

@@ -1,0 +1,19 @@
+#------------------------------------------------#
+# Configs for Laplacian Eigenmaps Figure 1.
+#------------------------------------------------#
+
+### General configs.
+D <- 10
+d <- 3
+# ns <- 2000
+ns <- round(seq(400,2000,length.out = 10))
+iters <- 50
+
+### Configs for sampling. ###
+sample_X <- make_sample_manifold(d,D)
+make_f0 <- make_spectral_sobolev_manifold
+
+
+### Methods. ###
+methods <- list(laplacian_eigenmaps = make_laplacian_eigenmaps)
+initialize_thetas <- list(initialize_laplacian_eigenmaps_thetas)

### 2021-05-21 code/plot_demos/eigenmaps_approximates_spectral_projection.R
Added scripts for creating plots for Laplacian Eigenmaps experiments.

@@ -1,0 +1,158 @@
+#--------------------------------------------------#
+# Script to generate plots used for Figure 2 of Laplacian Eigenmaps paper.
+#--------------------------------------------------#
+library(dplyr)
+library(ggplot2)
+library(gridExtra)
+library(viridis)
+source("plot_methods.R")
+
+# User entered information.
+data_directory <- "data/laplacian_eigenmaps/eigenmaps_parameters"
+plot_directory <- "plots/laplacian_eigenmaps/eigenmaps_approximates_spectral_projection" # Please change this to whichever directory you prefer.
+if(!exists(plot_directory)){dir.create(plot_directory)}
+
+# Load data
+load(file.path(data_directory,"configs.R"))
+d <- configs$d; ns <- configs$ns; methods <- configs$methods
+load(file.path(data_directory,"best_fits_by_method.R"))
+load(file.path(data_directory,"thetas.R"))
+load(file.path(data_directory,"mse.R"))
+load(file.path(data_directory,"Xs.R"))
+load(file.path(data_directory,"f0s.R"))
+load(file.path(data_directory,"Ys.R"))
+
+# Plotting parameters for all plots
+
+## Plot 1: Regression fxn and fitted values.
+plot1_indx <- 5
+plot1_K <- 10
+rs <- unique(thetas[[plot1_indx]][[1]]$r)
+plot1_r <- rs[which.min(abs(rs - .02))]    # arbitrary choice
+
+# Regression fxn
+pdf(file.path(plot_directory,"f0.pdf"))
+par(mar = c(5.1,6,4.1,4.1)) # Prevent the left hand side from getting cut off.
+plot_fxn(Xs[[plot1_indx]],Ys[[plot1_indx]],f = f0s[[plot1_indx]])
+dev.off()
+
+# Eigenmaps fitted values
+pdf(file.path(plot_directory,"eigenmaps.pdf"))
+par(mar = c(5.1,6,4.1,4.1)) # Prevent the left hand side from getting cut off.
+f <- subset(fits[[plot1_indx]][[1]], 
+            thetas[[plot1_indx]][[1]]$r == plot1_r & 
+            thetas[[plot1_indx]][[1]]$K == plot1_K)[1,]
[…]

### 2021-05-21 code/plot_demos/eigenmaps_mse.R

@@ -1,0 +1,36 @@
+#--------------------------------------------------#
+# Script to generate plots used for Figure 1 of Laplacian Eigenmaps paper.
+#--------------------------------------------------#
+library(dplyr)
+library(ggplot2)
+library(gridExtra)
+library(viridis)
+source("misc.R")
+source("plot_methods.R")
+
+# User entered information.
+data_directory <- "data/laplacian_eigenmaps/mse/eigenmaps_mse_2d_2s"
+plot_directory <- "plots/laplacian_eigenmaps/eigenmaps_mse" # Please change this to whichever directory you prefer.
+if(!exists(plot_directory)){dir.create(plot_directory)}
+
+load(file.path(data_directory,"configs.R"))
+d <- configs$d; ns <- configs$ns; s <- configs$s; methods <- configs$methods
+load(file.path(data_directory,"best_fits_by_method.R"))
+load(file.path(data_directory,"thetas.R"))
+load(file.path(data_directory,"mse.R"))
+load(file.path(data_directory,"Xs.R"))
+load(file.path(data_directory,"f0s.R"))
+load(file.path(data_directory,"Ys.R"))
+
+## Plot 1: Mean squared error as a function of n.
+plot_name <- paste0("mse_by_sample_size_",d,"d_",s,"s.pdf")
+pdf(file.path(plot_directory,plot_name))
+par(mar = c(5.1,6,4.1,4.1)) # Prevent the left hand side from getting cut off.
+
+## ALDEN GET RID OF
+best_K <- find_best_K(mse,thetas)
+best_radius <- find_best_radius(mse,thetas) 
+##
+
+plot_best_mse(methods,mse)
+dev.off()

### 2021-05-21 code/plot_demos/eigenmaps_tuning_parameters.R

@@ -1,0 +1,105 @@
+#--------------------------------------------------#
+# Script to generate plots analyzing impact of LE tuning parameters.
+#--------------------------------------------------#
+library(dplyr)
+library(ggplot2)
+library(gridExtra)
+library(viridis)
+source("plot_methods.R")
+
+# User entered information.
+data_directory <- "data/laplacian_eigenmaps/eigenmaps_parameters_2s"
+plot_directory <- "plots/laplacian_eigenmaps/eigenmaps_parameters_2s" # Please change this to whichever directory you prefer.
+if(!exists(plot_directory)){dir.create(plot_directory)}
+
+# Load data
+load(file.path(data_directory,"configs.R"))
+d <- configs$d; ns <- configs$ns; methods <- configs$methods
+load(file.path(data_directory,"best_fits_by_method.R"))
+load(file.path(data_directory,"thetas.R"))
+load(file.path(data_directory,"mse.R"))
+load(file.path(data_directory,"Xs.R"))
+load(file.path(data_directory,"f0s.R"))
+load(file.path(data_directory,"Ys.R"))
+
+if(!exists("s")) s <- 2 # ALDEN REMOVE THIS HACK
+
+# Plotting parameters for all plots
+title <- paste0("d = ",d,", s = ",s,".")
+
+## Plot 1: Mean squared error as a function of K.
+
+# Parameters
+alg_indx <- which(names(configs$methods) == "laplacian_eigenmaps")
+n_indx <- length(ns)
+all_rs <- unique(thetas[[n_indx]][[alg_indx]]$r)
+rs <- all_rs[round(seq(1,length(all_rs),length.out = 9))]
+Ks <- unique(thetas[[n_indx]][[alg_indx]][,"K"])
+
+# Data for plotting
+plot_mse <- mse[[n_indx]][[alg_indx]][which(thetas[[n_indx]][[alg_indx]]$r %in% rs),] %>%
+  rowMeans()
+plot_rs <- thetas[[n_indx]][[alg_indx]][which(thetas[[n_indx]][[alg_indx]]$r %in% rs),"r"]
+plot_Ks <- thetas[[n_indx]][[alg_indx]][which(thetas[[n_indx]][[alg_indx]]$r %in% rs),"K"]
[…]

### 2021-05-21 notes/misc/splines_graph_laplacians_connections.tex
Updated notes on connections between smoothing splines and graph Laplacian smoothing.

@@
 
-Suppose now that in addition to the \emph{labeled} data $(x_1,y_1),\ldots,(x_n,y_n)$, we observe $m$ \emph{unlabeled} design points $x_{n + 1},\ldots,x_{n + m}$ Let $N = n + m$, let $G_{N}$ denote the weighted graph over $x_1,\ldots,x_N$, and $L_{N,\varepsilon} \in \Reals^{N \times N}$ denote the graph Laplacian of $G_{N}$.
+Suppose now that in addition to the \emph{labeled} data $(x_1,y_1),\ldots,(x_n,y_n)$, we observe $m$ \emph{unlabeled} design points $x_{n + 1},\ldots,x_{n + m}$.
+Let $N = n + m$, let $G_{N}$ denote the weighted graph over $x_1,\ldots,x_N$, and $L_{N,\varepsilon} \in \Reals^{N \times N}$ denote the graph Laplacian of $G_{N}$.
 Thus $L_{N,\varepsilon}$ satisfies,for any $u \in \Reals^N$,
@@
 Finally, let $G_n^{(u)} := ([n], W_n^{(u)})$ be a weighted graph, with the $(i,j)$th entry of weighted edge matrix $W_n^{(u)}$ given by
-\begin{equation*}
+\begin{equation}
+\label{eqn:induced_insample_graph}
 \bigl(W_{n}^{(u)}\bigr)_{ij} := \sum_{k,\ell = m + 1}^{m + n} \eta\biggl(\frac{\|x_i - x_k\|}{\varepsilon}\biggr) \cdot \eta\biggl(\frac{\|x_j - x_{\ell}\|}{\varepsilon}\biggr) \cdot \Bigl(m\varepsilon^{d + 2}L_{m,\varepsilon} + D_{m,n}\Bigr)_{k,\ell}^{-1}.
-\end{equation*}
+\end{equation}
 Let $L_{n,\varepsilon}^{(u)}$ be the corresponding Laplacian, meaning
@@
 		\label{eqn:ssl_graph_laplacian_smoothing_one_unlabeled_1}
-		\wh{f}_N^{(\ell)} := \biggl(I_n + \frac{\rho n}{n + 1}\biggl[\frac{n}{n + 1}L_{n,\varepsilon} + \frac{1}{n + 1}L_{n,\varepsilon}^{(1)}\biggr]\biggr)^{-1} Y.
+		\wh{f}_N^{(\ell)} := \biggl(I_n + \frac{\rho n}{n + 1}\biggl[\frac{n}{n + 1}L_{n,\varepsilon} + \frac{1}{n + 1}L_{n,\varepsilon}^{(u)}\biggr]\biggr)^{-1} Y.
 		\end{equation}
@@
 
+Interestingly, the kernel smoothing step in~\eqref{eqn:ssl_graph_laplacian_smoothing_one_unlabeled_2} allows us to make dramatically stronger claims about the out-of-sample squared error $|\wh{a} - f_0(x_{n + 1})|^2$ than we could about the in-sample mean squared error $\|\wh{f}_n - f_0\|_n^2$.
+We will assume the following.
+\begin{assumption}[Design distribution $P$.]
+	\label{asmp:design2}
+	The design distribution $P$ is supported on an open, bounded set $\Omega \subseteq \Rd$, which has $C^{1}$ boundary $\partial \Omega$.
+The distribution $P$ admits a $C^{1}(\Omega)$ density $p$ with respect to Lebesgue measure $\nu$, which is bounded away from $0$ and $\infty$, meaning that for some $p_{\min} < 1 < p_{\max}$
+	\begin{equation*}
+	0 \leq p_{\min} \leq p(x) \leq p_{\max} < \infty,\quad\textrm{for all $x \in \Omega$.}
+	\end{equation*}
+\end{assumption}  
+\begin{theorem}
+	\label{thm:out_of_sample_laplacian_regularization}
+	Suppose $(x_1,y_1),\ldots,(x_n,y_n),x_{n + 1}$ are sampled from a distribution $P$ which satisfies Assumption~\ref{asmp:design2}, that the regression function $f_0$ is \textcolor{red}{$M$-Lipschitz} on $\Omega$.
+If the estimator $\wh{f}_N$ in~\eqref{eqn:ssl_graph_laplacian_smoothing_0} is computed with $0 < \rho \leq M^2(M^2n)^{-2/(2 + d)}$ and $\varepsilon = n^{-1/(2 + d)}$, then
+	\begin{equation*}
+	\bigl|\wh{a} - f_0(x_{n + 1})\bigr|^2 \leq C\frac{M^2}{\delta^2}(M^2n)^{-2/(2 + d)},
+	\end{equation*}
+	with probability at least $1 - C_1\delta - C_2\exp(-c_2n\varepsilon^d)$.
+\end{theorem}
+
 \section{Proofs}
@@
 
-\subsection{Proof of Theorem~{(?)}}
+\subsection{Proof of Theorem~\ref{thm:out_of_sample_laplacian_regularization}}
 \label{subsec:pf_ssl_graph_laplacian_smoothing_2}
@@
 It also suggests a route to analyzing the out-of-sample error $(\wh{a} - f_0(x_{n + 1}))^2$.
-Concretely, consider the following expansion
+Let $S_N$ and $S_n$ denote the smoother matrices such that $\wh{f}_N^{(\ell)} = S_NY$ and $\wh{f}_n = S_nY$; explicitly
+\begin{equation*}
+S_N := \Biggl(I_n + \frac{\rho n}{N}\biggl[\frac{n}{N}L_{n,\varepsilon} + \frac{1}{N}L_{n,\varepsilon}^{(u)}\biggr]\Biggr)^{-1}, \quad \textrm{and} \quad S_n := \Biggl(I_n + \frac{\rho n}{N}L_{n,\varepsilon}\Biggr)^{-1}.
+\end{equation*}
+Now we consider the following decomposition
 \begin{align*}
-\wh{a} - f_0(x_{n + 1}) & = \wh{a} - \frac{1}{d_{n,\varepsilon}(x_{n + 1})}K_{(x_{n + 1})} \Ebb_n\wh{f}_N^{(\ell)} + \\
-& \quad \frac{1}{d_{n,\varepsilon}(x_{n + 1})}K_{(x_{n + 1})} \Ebb_n\wh{f}_N^{(\ell)} - \frac{1}{d_{n,\varepsilon}(x_{n + 1})}K_{1,n} f_0 + \\
-& \quad \frac{1}{d_{n,\varepsilon}(x_{n + 1})}K_{(x_{n + 1})} f_0 - f_0(x_{n + 1}) \\
-\Longrightarrow \Ebb_n\bigl[|\wh{a} - f_0(x_{n + 1})|^2\bigr]& \leq \underbrace{\|T_{n,\varepsilon}(\wh{f}_N^{(\ell)} - \mathbb{E}_{n}\wh{f}_N^{(\ell)})\|_P^2}_{\textrm{variance}} + \underbrace{\|T_{n,\varepsilon}(\mathbb{E}_{n}\wh{f}_N^{(\ell)} - f_0)\|_P^2}_{\textrm{bias}} + \underbrace{\|T_{n,\varepsilon}f_0 - f_0\|_P^2}_{\textrm{noiseless kernel smoothing}}
+\wh{a} - f_0(x_{n + 1}) & = T_{n,\varepsilon}S_NY(x_{n + 1}) - T_{n,\varepsilon}S_Nf_0(x_{n + 1}) + T_{n,\varepsilon}S_Nf_0(x_{n + 1}) - T_{n,\varepsilon}S_nf_0(x_{n + 1}) + T_{n,\varepsilon}S_nf_0(x_{n + 1}) - T_{n,\varepsilon}f_0(x_{n + 1}) + \\
+& \quad T_{n,\varepsilon}f_0(x_{n + 1}) - f_0(x_{n + 1}) \\
+& =  \underbrace{T_{n,\varepsilon}S_Nw(x_{n + 1})}_{\textrm{Term 1}} + \underbrace{T_{n,\varepsilon}[(S_N - S_n)f_0](x_{n + 1})}_{\textrm{Term 2}} +  \underbrace{T_{n,\varepsilon}S_nf_0(x_{n + 1}) - T_{n,\varepsilon}f_0(x_{n + 1})}_{\textrm{Term 3}} + \underbrace{T_{n,\varepsilon}f_0(x_{n + 1}) - f_0(x_{n + 1})}_{\textrm{Term 4}}.
 \end{align*}
-Here $\Ebb_n$ denotes the expectation conditional on the labeled design points $x_1,\ldots,x_n$. Now we proceed to bound each of these terms, with high probability.
+The first of these terms is the contribution of the noise in $Y$.
+The second of these terms is the perturbation caused by the different smoother matrices.
+The third of these terms is the bias due to Laplacian regularization.
+The fourth term is the error inherent to kernel smoothing of noiseless samples.
+Now we proceed to upper bound the error induced by each of these terms.
 
-\paragraph{Term 1: Variance.}
-Let $\wt{L}_{n,\varepsilon} = n/(n + 1) L_{n,\varepsilon} + 1/(n + 1) L_{n,\varepsilon}^{(u)}$.
-Denoting the smoother matrix $S_n := [I_n + \rho n/(n + 1)\wt{L}_{n,\varepsilon}]^{-1}$, we can write the variance term as
+\paragraph{Term 1: Noise contribution.}
+Let $\Ebb_N$ denote the expectation conditional on the design points $x_1,\ldots,x_{n + 1}$. For any $i,j$ in $1,\ldots,n$, we have
 \begin{equation*}
-\|T_{n,\varepsilon}(\wh{f}_N^{(\ell)} - \mathbb{E}_{n}\wh{f}_N^{(\ell)})\|_P^2 =  \int_{\Omega} \Bigl(T_{n,\varepsilon}(S_nw)(x)\Bigr)^2 \,dP(x).
+\Ebb_N\bigl[(S_Nw)_i \cdot (S_nw)_j\bigr] = (S_N^2)_{ij}.
 \end{equation*}
-
-Note that for any $i,j$ in $1,\ldots,n$, we have
-\begin{equation*}
-\Ebb_n\bigl[(S_nw)_i \cdot (S_nw)_j\bigr] = (S_n^2)_{ij}.
-\end{equation*}
-Therefore, at a given $x \in \Omega$ it follows that
-\begin{equation*}
-\Ebb_n\Bigl[\bigl(T_{n,\varepsilon}S_nw(x)\bigr)^2\Bigr] = \frac{1}{d_{n,\varepsilon}(x)^2} K_{(x)}^{\top} S_n^2 K_{(x)} \leq \frac{1}{d_{n,\varepsilon}(x)^2} \|S_n\|_{\textrm{op}}^2 \cdot \|K_{(x)}\|_2^2 \leq \frac{1}{d_{n,\varepsilon}(x)^2} \cdot \|K_{(x)}\|_2^2.
-\end{equation*}
+Also, $\|S_N\|_{\textrm{op}} \leq 1$. 
+Therefore
+\begin{equation}
+\label{pf:out_of_sample_laplacian_regularization_0}
+\Ebb_N\Bigl[\bigl(T_{n,\varepsilon}S_Nw(x_{n + 1})\bigr)^2\Bigr] = \frac{1}{d_{n,\varepsilon}(x_{n + 1})^2} K_{(x_{n + 1})}^{\top} S_N^2 K_{(x_{n + 1})} \leq \frac{1}{d_{n,\varepsilon}(x_{n + 1})^2} \|S_N\|_{\textrm{op}}^2 \cdot \|K_{(x_{n + 1})}\|_2^2 \leq \frac{1}{d_{n,\varepsilon}(x_{n + 1})^2} \cdot \|K_{(x_{n + 1})}\|_2^2.
+\end{equation}
 Now by Lemma~\ref{lem:max_degree_deviation}, we have that for both $K(x) = \eta(x)$ and $K = \bigl(\eta(x)\bigr)^2$, with probability at least $1 - C_2\exp\bigl(-c_2 \delta^2 \cdot n \varepsilon^d\bigr)$,
@@
 \end{equation*}
-Integrating with respect to $P$, and exchanging expectation with integration thus gives that with probability at least $1 - C_2\exp\bigl(-c_2 \delta^2 \cdot n \varepsilon^d\bigr)$,
+Plugging this back in to~\eqref{pf:out_of_sample_laplacian_regularization_0} gives
 \begin{equation*}
-\|T_{n,\varepsilon}(\wh{f}_N^{(\ell)} - \mathbb{E}_{n}\wh{f}_N^{(\ell)})\|_P^2 \leq \biggl[\frac{(1 + \delta)}{(1 - \delta)}\biggr]^2 \frac{\|\eta\|_{\infty} p_{\max}}{n p_{\min} \varepsilon^d}.
+\Ebb_N\Bigl[\bigl(T_{n,\varepsilon}S_Nw(x_{n + 1})\bigr)^2\Bigr] \leq \biggl[\frac{(1 + \delta)}{(1 - \delta)}\biggr]^2 \frac{\|\eta\|_{\infty}}{n p_{\min} \varepsilon^d}.
+\end{equation*}
+Thus by Markov's inequality, with probability at least $1 - \beta - C_2\exp\bigl(-c_2 \delta^2 \cdot n \varepsilon^d\bigr)$
+\begin{equation*}
+\bigl(T_{n,\varepsilon}S_Nw(x_{n + 1})\bigr)^2 \leq \frac{1}{\beta}\biggl[\frac{(1 + \delta)}{(1 - \delta)}\biggr]^2 \frac{\|\eta\|_{\infty}}{n p_{\min} \varepsilon^d} = C_1\frac{1}{n\varepsilon^d}.
 \end{equation*}
 
-\paragraph{Term 2: Bias.}
-\textcolor{red}{(TODO)}.
+\paragraph{Term 2: Perturbation.}
+From the Cauchy-Schwarz inequality, we have that for any $u \in \Reals^n$,
+\begin{equation*}
+|T_{n,\varepsilon}u(x_{n + 1})|^2 = \biggl|\frac{1}{d_{n,\varepsilon}(x_{n + 1})} \sum_{i = 1}^{n} u_i \eta\biggl(\frac{\|x_i - x_{n + 1}\|}{\varepsilon}\biggr)\biggr|^2 \leq \frac{1}{d_{n,\varepsilon}(x_{n + 1})} \sum_{i = 1}^{n} |u_i|^2 \eta\biggl(\frac{\|x_{i} - x_{n + 1}\|}{\varepsilon}\biggr) \leq \|u\|_{\infty}^2.
+\end{equation*}
+Thus to upper bound Term 2 we focus our attention on upper bounding $\|(S_N - S_n)f_0\|_{\infty}$.
 
-\paragraph{Term 3: Noiseless kernel smoothing.}
-This is a standard analysis. \textcolor{red}{From our paper on Laplacian eigenmaps (equation (107) and Lemma~19)}, we have that
+Since $S_N$ and $S_n$ are symmetric, and thus $S_NS_n = S_nS_N$, it follows that
 \begin{equation*}
-\|T_{n,\varepsilon}f_0 - f_0\|_P^2 \leq \frac{C}{\delta} \cdot \biggl(\frac{\varepsilon^2}{n\varepsilon^d} + \varepsilon^2\biggr) |f|_{H^1(\Omega)}^2 \leq \frac{C}{\delta} \cdot \varepsilon^2 |f|_{H^1(\Omega)}^2
+(S_N - S_n)f_0 = (S_NS_nS_n^{-1} - S_nS_NS_N^{-1})f_0 = S_N S_n(S_n^{-1} - S_N^{-1})f_0.
 \end{equation*}
-with probability at least $1 - C\exp(-cn\varepsilon^d)$.
+Consequently we can upper bound the $\infty$-norm $\|(S_N - S_n)f_0\|_{\infty}$ by
+\begin{equation*}
+\|(S_N - S_n)f_0\|_{\infty} \leq \|S_N\|_{\infty} \|S_n\|_{\infty} \|(S_n^{-1} - S_N^{-1})f_0\|_{\infty}.
+\end{equation*}
+Here we recall that the $\infty$-operator norm of a matrix is $\|A\|_{\infty} = \max\{\|Au\|_{\infty}:\|u\|_{\infty} = 1\}$.
+
+We first upper bound $\|(S_n^{-1} - S_N^{-1})f_0\|_{\infty}$, using the decomposition
+\begin{equation*}
+S_n^{-1} - S_N^{-1} = \rho \biggl[\biggl(1 - \frac{n^2}{N^2}\biggr)L_{n,\varepsilon} - \frac{n}{N^2}L_{n,\varepsilon}^{(u)}\biggr],
+\end{equation*}
+and therefore
+\begin{equation}
+\label{pf:out_of_sample_laplacian_regularization_2}
+\|(S_n^{-1} - S_N^{-1})f_0\|_{\infty} \leq \frac{\rho}{n} \Bigl(\|L_{n,\varepsilon}f_0\|_{\infty} + \|L_{n,\varepsilon}^{(u)}f_0\|_{\infty} \Bigr).
+\end{equation}
+From H\"{o}lder's inequality and the Lipschitz property of $f_0$, we deduce both that
+\begin{align*}
+\|L_{n,\varepsilon}f_0\|_{\infty} & = \max_{i = 1,\ldots,n} \biggl| \frac{1}{n\varepsilon^{d + 2}}\sum_{j = 1}^{n} \bigl(f_0(x_i) - f_0(x_j)\bigr) \eta\biggl(\frac{\|x_i - x_j\|}{\varepsilon}\biggr) \biggr| \\
+& \leq \max_{i = 1,\ldots,n}  \frac{M}{n\varepsilon^{d + 1}}\sum_{j = 1}^{n} \eta\biggl(\frac{\|x_i - x_j\|}{\varepsilon}\biggr) \\
+& \leq M\frac{1}{\varepsilon^{d + 1}},
+\end{align*}
+and
+\begin{align*}
+\|L_{n,\varepsilon}^{(u)}f_0\|_{\infty} & = \max_{i = 1,\ldots,n} \biggl| \frac{1}{\varepsilon^{d + 2}} \sum_{j = 1}^{n} \bigl(f_0(x_i) - f_0(x_j)\bigr) W_{n,ij}^{(u)}\biggr| \\
+& \leq \max_{i = 1,\ldots,n} \frac{M}{\varepsilon^{d + 1}} \sum_{j = 1}^{n} |W_{n,ij}^{(u)}|\\
+& \leq \frac{M}{\varepsilon^{d + 1}} \|\eta\|_{\infty},
+\end{align*}
+where we recall the definition of $W_{n}^{(u)}$ from~\eqref{eqn:induced_insample_graph}. Plugging these upper bounds back into~\eqref{pf:out_of_sample_laplacian_regularization_2} gives
+\begin{equation*}
+\|(S_n^{-1} - S_N^{-1})f_0\|_{\infty} \leq 2 \|\eta\|_{\infty}\frac{\rho}{n\varepsilon^{d + 1}}M.
+\end{equation*}
+On the other hand, it is not hard to see that $S_N$ and $S_n$ are contractions in $L^{\infty}(\Reals^n)$, meaning that $\|S_N\|_{\infty},\|S_n\|_{\infty} \leq 1$.
+We establish that $S_n$ is a contraction, and the exact same reasoning will hold with respect to $S_N$ as well.
+For any $u \in \Reals^n$, assuming without loss of generality that $u_1 = \|u\|_{\infty}$, we have that
+\begin{equation*}
+(S_n^{-1}u)_1 = u_1 + (\rho L_{n,\varepsilon}u)_1 \geq u_1,
+\end{equation*}
+which implies that $\|S_n^{-1}u\|_{\infty} \geq \|u\|_{\infty}$. Thus for any $u \in \Reals^n$, we also have that
+\begin{equation*}
+\|u\|_{\infty} = \|S_n^{-1}S_nu\|_{\infty} \geq \|S_nu\|_{\infty},
+\end{equation*}
+demonstrating that $S_n$ is a contraction.
+
+Plugging these upper bounds back into~\eqref{pf:out_of_sample_laplacian_regularization_2}, we conclude that
+\begin{equation*}
+\bigl(T_{n,\varepsilon}(S_N - S_n)f_0(x_{n + 1})\bigr)^2 \leq \|(S_N - S_n)f_0\|_{\infty}^2 \leq 4 \|\eta\|_{\infty}^2\biggl(\frac{\rho}{n\varepsilon^{d + 1}}\biggr)^2M^2 =: C_2 \biggl(\frac{\rho}{n\varepsilon^{d + 1}}\biggr)^2M^2.
+\end{equation*}
+
+\paragraph{Term 3: Bias.}
+We begin by taking an expectation of Term 3 conditional on the labeled design points $x_1,\ldots,x_n$ (which we denote by $\Ebb_n$):
+\begin{equation*}
+\Ebb_n\Bigl[\Bigl(T_{n,\varepsilon}S_nf_0(x_{n + 1}) - T_{n,\varepsilon}f_0(x_{n + 1})\Bigr)^2\Bigr] = \|T_{n,\varepsilon}S_nf_0 - T_{n,\varepsilon}f_0\|_P^2.
+\end{equation*}
+We have show in \textcolor{red}{our paper on Laplacian eigenmaps (equation 106)} that for any $u \in \Reals^n$,
+\begin{equation*}
+\|T_{n,\varepsilon}u - T_{n,\varepsilon}f_0\|_P^2 \leq C \|u - f_0\|_n^2,
+\end{equation*}
+and that under the assumption~\ref{asmp:design}, for any Sobolev function $f \in H^1(\Omega)$, with probability at least $1 - \delta$
+\begin{equation*}
+\|S_nf_0 -f_n\|_n^2 \leq \frac{C\rho}{\delta}|f_0|_{H^1(\mc{X})}^2.
+\end{equation*}
+Combining these two results, we have that
+\begin{equation*}
+\Ebb_n\Bigl[\Bigl(T_{n,\varepsilon}S_nf_0(x_{n + 1}) - T_{n,\varepsilon}f_0(x_{n + 1})\Bigr)^2\Bigr] \leq \frac{C_3\rho}{\delta}|f_0|_{H^1(\Omega)}^2
+\end{equation*}                                                                          
+with probability at least $1 - \delta$. Thus it follows from Lemma~\ref{lem:stochastic_boundedness} that with probability at least $1 - 2\delta$,
+\begin{equation*}
+\Bigl(T_{n,\varepsilon}S_nf_0(x_{n + 1}) - T_{n,\varepsilon}f_0(x_{n + 1})\Bigr)^2 \leq \frac{C_3\rho}{\delta^2}|f_0|_{H^1(\Omega)}^2.
+\end{equation*}
+
+\paragraph{Term 4: Noiseless kernel smoothing.}
+\textcolor{red}{From our paper on Laplacian eigenmaps (equation (107) and Lemma~19)}, we have that
+\begin{equation*}
+\|T_{n,\varepsilon}f_0 - f_0\|_P^2 \leq \frac{C}{\delta} \cdot \biggl(\frac{\varepsilon^2}{n\varepsilon^d} + \varepsilon^2\biggr) |f|_{H^1(\Omega)}^2 \leq \frac{C_4}{\delta} \cdot \varepsilon^2 |f|_{H^1(\Omega)}^2
+\end{equation*}
+with probability at least $1 - \delta - C\exp(-cn\varepsilon^d)$.
 
@@
 
+\subsection{Stochastic Boundedness}
+The following Lemma helps us convert between bounds in conditional expectation and bounds in probability.
+\begin{lemma}
+	\label{lem:stochastic_boundedness}
+	Let $Z \geq 0$ and $X$ be random variables, and let $0 \leq E[Z|X] \leq a$ with probability at least $1 - \delta$. Then
+	\begin{equation*}
+	\Pbb\biggl(Z \geq \frac{a}{\delta}\biggr) \leq 2\delta.
+	\end{equation*}
+\end{lemma}
+\begin{proof}
+	By the law of iterated expectation, 
+	\begin{equation*}
+	\Pbb(Z \geq \frac{a}{\delta}) = \Ebb[\Pbb(Z \geq \frac{a}{\delta}|X)] = \Ebb[\Pbb(Z \geq \frac{a}{\delta}|X) \wedge 1].
+	\end{equation*}
+	Since $E[Z|X] \leq a$ with probability at least $1 - \delta$,
+	\begin{align*}
+	\Ebb[\Pbb(Z \geq \frac{a}{\delta}|X) \wedge 1] & = \Ebb\Bigl[\bigl(\Pbb(Z \geq \frac{a}{\delta}|X) \wedge 1\bigr) \cdot (\1\{E[Z|X] \leq a\} + \1\{E[Z|X] \geq a\})\Bigr] \\
+	& \leq \Ebb\Bigl[\bigl(\Pbb(Z \geq \frac{a}{\delta}|X) \cdot (\1\{E[Z|X] \leq a\} + \Ebb\Bigl[\1\{E[Z|X] \geq a\})\Bigr] \\
+	& \leq 2\delta,
+	\end{align*}
+	with the last upper bound following from (conditional) Markov's inequality.
+\end{proof}
+
 \end{document}

### 2021-05-21 manuscripts/laplacian_eigenmaps/appendix.tex

@@
 & \quad \frac{1}{\varepsilon^{d + 2}}\sum_{j = 1}^{s - 1} \frac{1}{j!} \int_{\mc{X}} \bigl(d_x^jf\bigr)(x' - x)  r_{x'}^{q}(x;p) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \quad  + \\
-& \quad \frac{1}{\varepsilon^{d + 2}} \int_{\mc{X}} r_{x'}^j(x;f) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \\
+& \quad \frac{1}{\varepsilon^{d + 2}} \int_{\mc{X}} r_{x'}^j(x;f) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x').
+\end{align*}
+Here we have adopted the convention that $\sum_{j = 1}^{0} = 0$.
+
+Changing variables to $z = (x' - x)/\varepsilon$, we can rewrite the above expression as 
+\begin{align*}
+L_{P,\varepsilon}f(x) & = \frac{1}{\varepsilon^{2}}\sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1}\frac{\varepsilon^{j_1 + j_2}}{j_1!j_2!}  \int d_x^{j_1}f(z) d_x^{j_2}p(z) \eta\bigl(\|z\|\bigr) \,dz \quad + \\
+& \quad \frac{1}{\varepsilon^{2}} \sum_{j = 1}^{s - 1} \frac{\varepsilon^j}{j!} \int d_x^jf(z)  r_{zh + x}^{q}(x;p) \eta\bigl(\|z\|\bigr) \,dz \quad  + \\
+& \quad \frac{1}{\varepsilon^{2}} \int r_{zh + x}^j(x;f) \eta\bigl(\|z\|\bigr) p(zh + x)\,dz \\
 & := G_1(x) + G_2(x) + G_3(x).
@@
 On the other hand, we will establish that if $s = 1$ then $G_2(x) = 0$, and otherwise for $s \geq 2$
-\begin{equation*}
+\begin{equation}
+\label{pf:approximation_error_nonlocal_laplacian_2}
 \|G_2\|_{\Leb^2(\mc{X}_{\varepsilon})} \leq C \varepsilon^{s - 2} M \|p\|_{C^{s - 1}(\mc{X})};
-\end{equation*}
+\end{equation}
 this same estimate will hold for $G_3$ for all $s \geq 1$. Together these will imply~\eqref{pf:approximation_error_nonlocal_laplacian_0} and~\eqref{pf:approximation_error_nonlocal_laplacian_1}.
@@
 \emph{Estimate on $G_1(x)$.}
-If $s = 1$, then $s - 1 = 0$, and so $G_1(x) = 0$. We may therefore suppose $s \geq 2$. In this case, a change of variables gives
-\begin{align}
-L_{P,\varepsilon}f(x) & = \frac{1}{\varepsilon^{d + 2}} \sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1} \frac{1}{j_1!j_2!}  \int_{\mc{X}} \bigl(d_x^{j_1}f\bigr)(x' - x) \bigl(d_x^{j_2}p\bigr)(x' - x) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \nonumber \\
-& = \sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1} \frac{\varepsilon^{j_1 + j_2 - 2}}{j_1!j_2!}  \underbrace{\int_{B(0,1)} d_x^{j_1}f(z) d_x^{j_2}p(z) \eta(\|z\|) \,dz}_{:= g_{j_1,j_2}(x)} \label{pf:approximation_error_nonlocal_laplacian_3}
-\end{align}
+If $s = 1$, then $s - 1 = 0$, and so $G_1(x) = 0$. We may therefore suppose $s \geq 2$. Recall that
+\begin{equation}
+G_1(x) = \sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1} \frac{\varepsilon^{j_1 + j_2 - 2}}{j_1!j_2!}  \underbrace{\int_{B(0,1)} d_x^{j_1}f(z) d_x^{j_2}p(z) \eta(\|z\|) \,dz}_{:= g_{j_1,j_2}(x)} \label{pf:approximation_error_nonlocal_laplacian_3}
+\end{equation}
 The nature of $g_{j_1,j_2}(x)$ depends on the sum $j_1 + j_2$.
@@
 \end{equation*}
-In particular this is the case when $j_1 = 1$ and $j_2 = 0$, which means that if $s = 2$, then $G_1(x) = 0$.
-On the other hand if $s \geq 3$, then the lowest order term in~\eqref{pf:approximation_error_nonlocal_laplacian_3} is when $j_1 + j_2 = 2$, so that either $j_1 = 1$ and $j_2 = 1$, or $j_1 = 2$ and $j_2 = 0$.
-It follows that
+In particular this is the case when $j_1 = 1$ and $j_2 = 0$.
+Thus when $s = 2$,  $G_1(x) = g_{1,0}(x) = 0$.
+On the other hand if $s \geq 3$, then the lowest order terms in~\eqref{pf:approximation_error_nonlocal_laplacian_3} are those where $j_1 + j_2 = 2$, so that either $j_1 = 1$ and $j_2 = 1$, or $j_1 = 2$ and $j_2 = 0$.
+We have that
 \begin{align*}
@@
 & = \sum_{i = 1}^{d} D^{e_{i}}f(x) D^{e_{i}}p(x) \int_{\mc{X}} z^2 \eta(\|z\|) \,dz + \frac{p(x)}{2} \sum_{i = 1}^{d} D^{2e_{i}}f(x)\int_{\mc{X}} z^2 \eta(\|z\|) \,dz\\ 
-& = \sigma_{\eta}\Delta_Pf(x).
-\end{align*}
+& = \sigma_{\eta}\Delta_Pf(x),
+\end{align*}
+which is the leading term order term.
 Now it remains only to deal with the higher-order terms, where $j_1 + j_2 > 2$, and where it suffices to show that each function $g_{j_1,j_2}$ satisfies~\eqref{pf:approximation_error_nonlocal_laplacian_1.5} for $j = \min\{j_1 + j_2 - 2,s - 2\}$.
-Now it is helpful to write $g_{j_1,j_2}$ using multi-index notation,
+It is helpful to write $g_{j_1,j_2}$ using multi-index notation,
 \begin{align*}
@@
 \begin{align}
-|G_2(x)| & \leq \frac{1}{\varepsilon^{d + 2}}\sum_{j = 1}^{s - 1} \frac{1}{j!}\int_{\mc{X}} \Bigl|\bigl(d_x^{j}f\bigr)(x' - x)\Bigr| \cdot \Bigl|r_{x'}^{s - 1}(x;p)\Bigr| \cdot \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \nonumber \\
-& = \frac{1}{\varepsilon^2} \sum_{j = 1}^{s - 1} \frac{1}{j!}\int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(\varepsilon z)\Bigr| \cdot |r_{x + z\varepsilon}^{s - 1}(x;p)| \cdot \eta(\|z\|) \,dz \nonumber \\
+|G_2(x)| & \leq \sum_{j = 1}^{s - 1} \frac{\varepsilon^{j - 2}}{j!}\int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(z)\Bigr| \cdot |r_{x + z\varepsilon}^{s - 1}(x;p)| \cdot \eta(\|z\|) \,dz \nonumber \\
 & \leq C\varepsilon^{s - 2}\|p\|_{C^{s - 1}(\mc{X})} \sum_{j = 1}^{s - 1} \frac{1}{j!} \int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(z)\Bigr| \cdot \eta(\|z\|) \,dz \label{pf:approximation_error_nonlocal_laplacian_4}.
@@
 \emph{Estimate on $G_3(x)$.}
-We begin with a pointwise upper bound on $|J_3(x)|$, 
-\begin{align*}
-|G_3(x)| & \leq \frac{p_{\max}}{\varepsilon^{2 + d}} \int_{\mc{X}} \bigl|r_{x'}^s(x;f)\bigr| \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \\
-& = \frac{p_{\max}}{\varepsilon^2} \int_{B(0,1)} \bigl|r_{x + \varepsilon z}^s(x;f)\bigr| \eta(\|z\|) \,dz,
-\end{align*}
-and applying the Cauchy-Schwarz inequality, we deduce a pointwise upper bound on $|G_3(x)|^2$,
+Applying the Cauchy-Schwarz inequality, we deduce a pointwise upper bound on $|G_3(x)|^2$,
 \begin{align*}
@@
 \begin{equation*}
-A_1(x) \leq \frac{8L_{\eta}\varepsilon^2}{R^2}\int_{\mc{X}}\bigl[f(x') - f(x)\bigr]^2 \1\{\|x' - x\| \leq \varepsilon\} \,d\mu(x') + \|\eta\|_{\infty} \int_{\mc{X}}\bigl[f(x') - f(x)\bigr]^2 \1\bigl\{\varepsilon \|x' - x\| \leq \wt{\varepsilon}\bigr\} \,d\mu(x') \\
+A_1(x) \leq \frac{8L_{\eta}\varepsilon^2}{R^2}\int_{\mc{X}}\bigl[f(x') - f(x)\bigr]^2 \1\{\|x' - x\| \leq \varepsilon\} \,d\mu(x') + \|\eta\|_{\infty} \int_{\mc{X}}\bigl[f(x') - f(x)\bigr]^2 \1\bigl\{\varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon}\bigr\} \,d\mu(x') \\
 \end{equation*}
@@
 	Suppose Model~\ref{def:model_manifold}, and additionally that $f \in H^1(\mc{X})$. There exist constants $c$ and $C$ such that for any $\varepsilon < c$,
-	\begin{equation*}
-	\int_{\mc{X}} \int_{\mc{X}} \bigl(f(x') - f(x)\bigr)^2 \1\{\varepsilon < \|x' - x\| < \wt{\varepsilon}\} \,dP(x') \,dP(x) \leq C \varepsilon^{4 + m} \|f\|_{H^1(\mc{X})}^2
-	\end{equation*}
+	\begin{equation}
+	\label{eqn:dirichlet_energy_remainder}
+	\int_{\mc{X}} \int_{\mc{X}} \bigl(f(x') - f(x)\bigr)^2 \1\{\varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon}\} \,d\mu(x') \,d\mu(x) \leq C \varepsilon^{4 + m} \|f\|_{H^1(\mc{X})}^2
+	\end{equation}
 \end{lemma}
@@
 \end{equation*}
-From the lower bound in~\eqref{eqn:distance_error}, it follows that $E_{P,\varepsilon}(f;X) \leq \wt{E}_{P,\varepsilon}(f;{\mc{X}})$, and from the upper bounds $p(x) \leq p_{\max}$ and $\eta(|x|) \leq \|\eta\|_{\infty} \cdot \1{x \in [-1,1]}$ we further have
+From the lower bound in~\eqref{eqn:distance_error}, it follows that $E_{P,\varepsilon}(f;X) \leq \wt{E}_{P,\varepsilon}(f;{\mc{X}})$, and from the upper bounds $p(x) \leq p_{\max}$ and $\eta(|x|) \leq \|\eta\|_{\infty} \cdot \1\{x \in [-1,1]\}$ we further have
 \begin{equation*}
@@
 \paragraph{Proof (of Lemma~\ref{lem:dirichlet_energy_remainder}).}
-\textcolor{red}{(TODO)}
+Following exactly the steps of the proof of Lemma~3.3 of \citet{burago2014}, but replacing all references to a ball of radius $r$ by references to the set difference between balls of radius $\wt{\varepsilon}$ and $\varepsilon$, we obtain that
+\begin{equation*}
+\int_{\mc{X}} \int_{\mc{X}} \bigl(f(x') - f(x)\bigr)^2 \1\{\varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon}\} \,d\mu(x') \,d\mu(x) \leq (1 + CmK\varepsilon^2) \cdot \int_{\mc{X}} \int_{B_{m}(0,\wt{\varepsilon})} |d_x^{1}f(v)|^2 \,dv \,d\mu(x).
+\end{equation*}
+From (2.7) of~\citet{burago2014}, we further have
+\begin{equation*}
+\int_{\mc{X}} \int_{B_{m}(0,\wt{\varepsilon})} |d_x^{1}f(v)|^2 \,dv \,d\mu(x)  = \frac{\nu_m}{2 + m} (\wt{\varepsilon}^{2 + m} - \varepsilon^{2 + m}) \int_{\mc{X}} |d_x^1f|^2 \,d\mu(x) = 27\frac{\nu_m}{(2 + m)R^2} \varepsilon^{4 + m} \|d^1f\|_{\Leb^2(\mc{X})}^2.
+\end{equation*}
+Recalling that $\|d^1f\|_{\Leb^2(\mc{X})}^2 \leq \|f\|_{H^1(\mc{X})}^2$, we see that this implies the claim of Lemma~\ref{lem:dirichlet_energy_remainder}.
 
@@
 In this section we prove Lemma~\ref{lem:kernel_smoothing_insample} (in Section~\ref{subsec:pf_kernel_smoothing_insample}) and Lemma~\ref{lem:kernel_smoothing_bias} (in Section~\ref{subsec:pf_kernel_smoothing_bias}).
-In Section~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we give a sequence of design densities and regression functions $\{p^{(n)}(x), f_0^{(n)}(x)\}_{n \in \mathbb{N}}$ for which the estimator $T_{h,n}\wc{f}$ (kernel extension of Laplacian eigenmaps) strictly outperforms directly kernel smoothing the responses, in the sense that
-\begin{equation*}
-\lim_{n \to \infty} \sup_{h'} \frac{\Ebb \|T_{h,n}\wh{f} - f_0\|_P^2}{\Ebb \|T_{h',n}Y - f_0\|_P^2 } = 0.
-\end{equation*}
+In Section~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we give a sequence of design densities and regression functions $\{p^{(n)}(x), f_0^{(n)}(x)\}_{n \in \mathbb{N}}$ for which the estimator $T_{n,h}\wc{f}$ (extension of Laplacian eigenmaps by kernel smoothing) strictly outperforms directly kernel smoothing the responses, in the sense that
+\begin{equation*}
+\lim_{n \to \infty} \sup_{h'} \frac{\Ebb \|T_{n,h}\wh{f} - f_0\|_P^2}{\Ebb \|T_{h',n}Y - f_0\|_P^2 } = 0.
+\end{equation*}
+We begin with some preliminary estimates in Section~\ref{subsec:kernel_smoothing_preliminaries}, which will ease the subsequent analysis.
+\subsection{Some preliminary estimates}
+\label{subsec:kernel_smoothing_preliminaries}
+In certain parts the analysis of this section will overlap with Section~\ref{sec:graph_quadratic_form_euclidean}, where we upper bounded the non-local graph-Sobolev seminorm of a function $f$ in terms of the Sobolev norm of $f$.
+To see why this should be, note that for an function $f$ and point $x \in \mc{X}$, we have
+\begin{equation*}
+T_{P,h}f(x) - f(x) = \frac{1}{d_{Q,h}(x)} \int \bigl(f(x') - f(x)\bigr) \psi\biggl(\frac{\|x' - x\|}{h}\biggr) \,dQ(x') = \frac{h^{d + 2}}{d_{P,h}(x)} L_{P,h}f(x).
+\end{equation*}
+This expression reflects the known fact that the bias operator of kernel smoothing is equal to the non-local Laplacian, up to a rescaling by the population degree functional $d_{P,h}(x)$.
+In the second equality, we are using the notation $L_{P,h}f(x)$ exactly as defined in~\eqref{eqn:nonlocal_laplacian}, but with the kernel $\psi$ instead of $\eta$.
+Note that $\psi$ satisfies all the same assumptions as $\eta$, except that of positivity; when $\psi$ is a higher-order kernel it may take negative values.
+
+Now we provide a lower bound on $d_{P,h}(x)$ that holds uniformly over all $x \in \mc{X}$.
+Recall that by assumption the density $p$ is Lipschitz.
+Letting $L_p$ denote the Lipschitz constant of $p$, we have that
+\begin{align*}
+d_{P,h}(x) & = \int \psi\biggl(\frac{\|x' - x\|}{h}\biggr) p(x') \,dx' \\
+& = h^d \int \psi(\|z\|) p(hz + x) \1\Bigl\{ hz + x \in \mc{X} \Bigr\} \,dz \\
+& \geq h^d p(x) \int \psi(\|z\|) \1\Bigl\{ hz + x  \in \mc{X}\Bigr\} \,dz - L_p h^{d + 1} \|\psi\|_{\infty} \nu_d.
+\end{align*}
+Since by assumption $\mc{X}$ has Lipschitz boundary, setting $c_0$ to be a sufficiently small constant in~\ref{asmp:bandwidth}, we can further deduce that $\int \psi(\|z\|) \1\{hz + x \in \mc{X}\} \,dz \geq 1/3$, and consequently that
+\begin{equation}
+\label{eqn:degree_lower_bound}
+d_{P,h}(x) \geq \frac{p(x)}{3} h^d \geq \frac{p_{\min}}{3}h^d \quad \textrm{for all $x \in \mc{X}$.}
+\end{equation}
 
@@
 
-To begin with, we apply the triangle inequality to upper bound $\|T_{h,n}\wc{f} - f_0\|_P$ by the sum of two terms,
+To begin with, we apply the triangle inequality to upper bound $\|T_{n,h}\wc{f} - f_0\|_P$ by the sum of two terms,
 \begin{equation}
 \label{pf:kernel_smoothing_insample_1}
-\|T_{h,n}\wc{f} - f_0\|_P \leq \|T_{h,n}(\wc{f} - f_0)\|_P + \|T_{h,n}f_0 - f_0\|_P.
+\|T_{n,h}\wc{f} - f_0\|_P \leq \|T_{n,h}(\wc{f} - f_0)\|_P + \|T_{n,h}f_0 - f_0\|_P.
 \end{equation}
@@
 \label{pf:kernel_smoothing_insample_2}
-\|T_{h,n}(\wc{f} - f_0)\|_P^2 \leq C \|\wc{f} - f_0\|_n^2
+\|T_{n,h}(\wc{f} - f_0)\|_P^2 \leq C \|\wc{f} - f_0\|_n^2
 \end{equation}
@@
 \label{pf:kernel_smoothing_insample_3}
-\|T_{h,n}f_0 - f_0\|_P^2 \leq \frac{C}{\delta} \cdot \frac{h^2}{nh^d} |f|_{H^1(\mc{X})}^2 + \frac{C}{\delta}\|T_{h,P}f_0 - f_0\|_P^2,
+\|T_{n,h}f_0 - f_0\|_P^2 \leq \frac{C}{\delta} \cdot \frac{h^2}{nh^d} |f|_{H^1(\mc{X})}^2 + \frac{C}{\delta}\|T_{h,P}f_0 - f_0\|_P^2,
 \end{equation}
@@
 \begin{align*}
-\Bigl[T_{h,n}\bigl(\wc{f} - f_0\bigr)(x)\Bigr]^2 & = \Biggl[\frac{1}{d_{n,h}(x)^2}\int \psi\biggl(\frac{\|x' - x\|}{h}\biggr) \cdot \bigl(\wc{f}(x') - f_0(x')\bigr) \,dP_n(x')\Biggr]^2 \\
+\Bigl[T_{n,h}\bigl(\wc{f} - f_0\bigr)(x)\Bigr]^2 & = \Biggl[\frac{1}{d_{n,h}(x)^2}\int \psi\biggl(\frac{\|x' - x\|}{h}\biggr) \cdot \bigl(\wc{f}(x') - f_0(x')\bigr) \,dP_n(x')\Biggr]^2 \\
 & \leq \Biggl[\frac{1}{d_{n,h(x)}^2} \int \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr| \,dP_n(x')\Biggr] \cdot \Biggl[\int \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr| \cdot \bigl(\wc{f}(x') - f_0(x')\bigr)^2 \,dP_n(x')\Biggr] \\
@@
 \begin{align}
-\|T_{h,n}(\wc{f} - f_0)\|_P^2 & = \int \Bigl[T_{h,n}\bigl(\wc{f} - f_0\bigr)(x)\Bigr]^2 \,dP(x) \nonumber \\
+\|T_{n,h}(\wc{f} - f_0)\|_P^2 & = \int \Bigl[T_{n,h}\bigl(\wc{f} - f_0\bigr)(x)\Bigr]^2 \,dP(x) \nonumber \\
 & \leq \int \int \frac{d_{n,h}^{+}(x)}{|d_{n,h}(x)|} \cdot \frac{1}{|d_{n,h}(x)|} \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr| \cdot \bigl(\wc{f}(x') - f_0(x')\bigr)^2 \,dP_n(x') \,dP(x) \nonumber \\
@@
 & \leq \| T_{P,h}f_0 - f_0\|_P^2 + \frac{1}{n}\int \int \frac{1}{\bigl(d_{P,h}(x)\bigr)^2} \bigl(f_0(x') - f_0(x)\bigr)^2 \cdot \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr|^2 \,dP(x') \,dP(x) \\
-& \leq \| T_{P,h}f_0 - f_0\|_P^2 + \frac{3}{p_{\min}nh^d} \wt{E}_{P,h}(f_0),
-\end{align*}
-where in the final inequality we have used the lower bound on $d_{P,h}(x)$ from~\eqref{eqn:degree_lower_bound}, and defined $\wt{E}_{P,h}(f_0)$ to be the non-local Dirichlet energy with respect to the kernel $\psi^2$.
+& \leq \| T_{P,h}f_0 - f_0\|_P^2 + \frac{3h^2}{p_{\min}n} \wt{E}_{P,h}(f_0;\psi^2).
+\end{align*}
+In the final inequality we have used the lower bound on $d_{P,h}(x)$ from~\eqref{eqn:degree_lower_bound}, and written $E_{P,h}(f_0;\psi^2)$ for the non-local Dirichlet energy defined with respect to the kernel $\psi^2$.
 
@@
 & \leq \sup_{x \in \mc{X}} \biggl[\frac{d_{P,h}(x)}{d_{n,h}(x)}\biggr]^2 \cdot \int \Bigl(\wt{L}_{n,h}f_0(x)\Bigr)^2 \,dP(x) \\
-& \overset{(i)}{\leq} 4\frac{\| T_{P,h}f_0 - f_0\|_P^2}{\delta} + \frac{12}{\delta p_{\min}nh^d} \wt{E}_{P,h}(f_0) \\
-& \overset{(ii)}{\leq}  4\frac{\| T_{P,h}f_0 - f_0\|_P^2}{\delta} + \frac{C}{\delta p_{\min}nh^d} h^2 |f_0|_{H^1(\mc{X})}^2
+& \overset{(i)}{\leq} 4\frac{\| T_{P,h}f_0 - f_0\|_P^2}{\delta} + \frac{12h^2}{\delta p_{\min}nh^d} {E}_{P,h}(f_0;\psi^2) \\
+& \overset{(ii)}{\leq}  4\frac{\| T_{P,h}f_0 - f_0\|_P^2}{\delta} + \frac{Ch^2}{\delta p_{\min}nh^d} |f_0|_{H^1(\mc{X})}^2,
 \end{align*}
 with probability at least $1 - \delta - C\exp(-cnh^d)$.
-In $(i)$ we have used Markov's inequality, and in $(ii)$ we have applied the estimate~\eqref{pf:estimate_nonlocal_seminorm_1} to the non-local Dirichlet energy $\wt{E}_{P,h}(f_0)$.
+In $(i)$ we have used Markov's inequality, and in $(ii)$ we have applied the estimate~\eqref{pf:estimate_nonlocal_seminorm_1} to the non-local Dirichlet energy ${E}_{P,h}(f_0;\psi^2)$.
 This establishes~\eqref{pf:kernel_smoothing_insample_3}.
@@
 \label{subsec:pf_kernel_smoothing_bias}
-
+Lemma~\ref{lem:kernel_smoothing_bias} gives the necessary upper bounds on the bias of kernel smoothing.
 \begin{lemma}
 	\label{lem:kernel_smoothing_bias}
-	Suppose Model~\ref{def:model_flat_euclidean}, and that the smoothing operator $T_{h,P}$ is computed with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel}.
+	Suppose Model~\ref{def:model_flat_euclidean}, and that the kernel smoothing operator $T_{P,h}$ is computed with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel}.
 	\begin{itemize}
-		\item If $f_0 \in H^1(\mc{X})$, it follows that
-		\begin{equation*}
-		\|T_{h,P}f_0 - f_0\|_P^2 \leq C h^{2} |f|_{H^1(\mc{X})}^2.
-		\end{equation*}
-		\item If $f_0 \in H_0^{s}(\mc{X})$, $p \in C^{s - 1}(\mc{X})$, it follows that
-		\begin{equation*}
-		\|T_{h,P}f_0 - f_0\|_P^2 \leq C h^{2s} |f|_{H^s(\mc{X})}^2.
+		\item If $f_0 \in H^1(\mc{X})$, then there exists a constant $C$ which does not depend on $f_0$ such that
+		\begin{equation*}
+		\|T_{P,h}f_0 - f_0\|_P^2 \leq C h^{2} |f|_{H^1(\mc{X})}^2.
+		\end{equation*}
+		\item If $f_0 \in H_0^{s}(\mc{X})$, $p \in C^{s - 1}(\mc{X})$, and $\eta$ satisfies~\ref{asmp:ho_kernel}, then there exists a constant $C$ which does not depend on $f_0$ such that
+		\begin{equation*}
+		\|T_{P,h}f_0 - f_0\|_P^2 \leq C h^{2s} |f|_{H^s(\mc{X})}^2.
 		\end{equation*}
@@
 \end{lemma}
+We separately prove the first-order ($s = 1$) and higher-order ($s > 1$) parts of Lemma~\ref{lem:kernel_smoothing_bias}.
+In both cases, the proof will rely heavily on results already established regarding the non-local Laplacian $L_{P,h}$ and non-local Dirichlet energy $E_{P,h}$, which we recall are given for a kernel function $\mc{K}$ by
+\begin{equation*}
+L_{P,h}f(x) = \frac{1}{h^{d + 2}} \int \bigl(f(x') - f(x)\bigr)\mc{K}\biggl(\frac{\|x' - x\|}{h}\biggr)\,dP(x'),
+\end{equation*}
+and $E_{P,h}(f;\mc{K}) = \dotp{L_{P,h}f}{f}_{P}$, respectively.
+
+\paragraph{Proof of Lemma~\ref{lem:kernel_smoothing_bias}, $s = 1$.}
+Using the conclusions from Section~\ref{subsec:kernel_smoothing_preliminaries}, we have that
+\begin{equation}
+\label{pf:kernel_smoothing_bias_1}
+\|T_{P,h}f - f\|_P^2 \leq \frac{9h^4}{p_{\min}^2} \int \bigl[L_{P,h}f(x)\bigr]^2 \,dP(x).
+\end{equation}
+By the Cauchy-Schwarz inequality, we have that
+\begin{align*}
+\int \bigl[L_{P,h}f(x)\bigr]^2 \,dP(x) & = \frac{1}{h^{2d + 4}}\int \biggl[\int \bigl(f(x') - f(x)\bigr)\psi\biggl(\frac{\|x' - x\|}{h}\biggr) \,dP(x')\biggr]^2 \,dP(x) \\
+& \leq \frac{C}{h^{d + 4}} \int \int \bigl(f(x') - f(x)\bigr)^2 \cdot \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr| \,dP(x') \,dP(x) \\
+& = \frac{C}{h^{2}} E_{P,h}(f;|\psi|).
+\end{align*}
+Applying the estimate~\eqref{pf:estimate_nonlocal_seminorm_1} to the non-local Dirichlet energy ${E}_{P,h}(f;|\psi|)$ and plugging back into~\eqref{pf:kernel_smoothing_bias_1} gives the claimed result.
+
+\paragraph{Proof of Lemma~\ref{lem:kernel_smoothing_bias}, $s > 1$.}
+Proceeding from~\eqref{pf:kernel_smoothing_bias_1}, we separate the integral into the portion sufficiently in the interior of $\mc{X}$ and that near the boundary, obtaining
+\begin{equation}
+\label{pf:kernel_smoothing_bias_2}
+\|T_{P,h}f - f\|_P^2 \leq\frac{9p_{\max}h^4}{p_{\min}^2}\Bigl(\|L_{P,h}f\|_{L^2(\mc{X}_h)}^2 + \|L_{P,h}f\|_{L^2(\partial_{h}(\mc{X}))}^2\Bigr).
+\end{equation}
+In Lemma~\ref{lem:approximation_error_nonlocal_laplacian_boundary}, we established a sufficient upper bound on the second term,
+\begin{equation*}
+\|L_{P,h}f\|_{L^2(\partial_{h}(\mc{X}))}^2 \leq Ch^{2(s - 2)} \|f\|_{H^s(\mc{X})}^2.
+\end{equation*}
+Thus it remains to upper bound the first term. Here we recall that at a given $x \in \mc{X}_h$, we can write
+\begin{align*}
+L_{P,h}f(x) & = \frac{1}{h^{2}}\sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1}\frac{h^{j_1 + j_2}}{j_1!j_2!}  \int d_x^{j_1}f(z) d_x^{j_2}p(z) \psi\bigl(\|z\|\bigr) \,dz \quad + \\
+& \quad \frac{1}{h^{2}} \sum_{j = 1}^{s - 1} \frac{h^j}{j!} \int d_x^jf(z)  r_{zh + x}^{q}(x;p) \psi\bigl(\|z\|\bigr) \,dz \quad  + \\
+& \quad \frac{1}{h^{2}} \int r_{zh + x}^j(x;f) \psi\bigl(\|z\|\bigr) p(zh + x)\,dz \\
+& = G_1(x) + G_2(x) + G_3(x).
+\end{align*}
+(Here $q = s - 1$.)
+
+We have already given sufficient upper bounds on $\|G_j\|_{L^2(\mc{X}_{h})}$ for $j = 2,3$ in~\eqref{pf:approximation_error_nonlocal_laplacian_2}.
+Thus it remains only to upper bound $\|G_1\|_{L^2(\mc{X}_h)}$.
+Recall the expansion of $G_1$ from~\eqref{pf:approximation_error_nonlocal_laplacian_3},
+\begin{equation*}
+G_1(x) = \sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1} \frac{h^{j_1 + j_2 - 2}}{j_1!j_2!}  \underbrace{\int_{B(0,1)} d_x^{j_1}f(z) d_x^{j_2}p(z) \eta(\|z\|) \,dz}_{:= g_{j_1,j_2}(x)}.
+\end{equation*}
+Noting that $d_x^{j_1} \cdot d_x^{j_2}$ is a degree-$(j_1 + j_2)$ multivariate polynomial, and recalling that $\psi$ is an order-$s$ kernel, we have that 
+\begin{equation*}
+\int g_{j_1,j_2}(z) \psi\bigl(\|z\|\bigr) \,dz = 0,\quad \textrm{for all $j_1,j_2$ such that $j_1 + j_2 < s$.}
+\end{equation*}
+Otherwise, derivations similar to those used in the proof of Lemma~\ref{lem:approximation_error_nonlocal_laplacian} imply that
+\begin{equation*}
+\|g_{j_1,j_2}\|_{L^2(\mc{X}_h)} \leq C \|f\|_{H^s(\mc{X})} \|p\|_{C^{s - 1}(\mc{X})},\quad \textrm{for all $j_1,j_2$ such that $j_1 + j_2 \geq s$,}
+\end{equation*}
+from which it follows that
+\begin{equation*}
+\|G_1\|_{L^2(\mc{X}_h)}^2 \leq C  h^{2(s - 2)} \|f\|_{H^s(\mc{X})} \|p\|_{C^{s - 1}(\mc{X})}.
+\end{equation*}
+Together these upper bounds on $\|G_j\|_{L^2(\mc{X}_j)}$ for $j = 1,2,3$ imply that
+\begin{equation*}
+\|L_{P,h}f\|_{\mc{X}_h}^2 \leq Ch^{2(s - 2)} \|f\|_{H^s(\mc{X})}^2,
+\end{equation*}
+and plugging this back into~\eqref{pf:kernel_smoothing_bias_2} yields the claim.
 
@@
 
-In this section, we give a sequence of densities $p^{(n)}$ and regression functions $f_0^{(n)}$ for which the $L^2(P)$ error of the two-stage estimator $T_{h,n}\wh{f}$ is stochastically dominated by the $L^2(P)$ error of applying kernel smoothing directly to the responses.
+In this section, we give a sequence of densities $p^{(n)}$ and regression functions $f_0^{(n)}$ for which the $L^2(P)$ error of the two-stage estimator $T_{n,h}\wh{f}$ is stochastically dominated by the $L^2(P)$ error of applying kernel smoothing directly to the responses.
 We begin by precisely setting up the problem and reviewing the two estimators we will compare.
@@
 Now we translate Lemma~\ref{lem:uniform_bound_empirical_degree} into a multiplicative bound, which will be more useful for our purposes. 
-By assumption, the density $p$ is Lipschitz.
-Letting $L_p$ denote the Lipschitz constant of $p$, we have that
-\begin{align*}
-d_{P,h}(x) & = \int \psi\biggl(\frac{\|x' - x\|}{h}\biggr) p(x') \,dx' \\
-& = h^d \int \psi(\|z\|) p(hz + x) \1\Bigl\{ hz + x \in \mc{X} \Bigr\} \,dz \\
-& \geq h^d p(x) \int \psi(\|z\|) \1\Bigl\{ hz + x  \in \mc{X}\Bigr\} \,dz - L_p h^{d + 1} \|\psi\|_{\infty} \nu_d.
-\end{align*}
-Since by assumption $\mc{X}$ has Lipschitz boundary, setting $c_0$ to be a sufficiently small constant in~\ref{asmp:bandwidth}, we can further deduce that $\int \psi(\|z\|) \1\{hz + x \in \mc{X}\} \,dz \geq 1/3$, and consequently that
-\begin{equation}
-\label{eqn:degree_lower_bound}
-d_{P,h}(x) \geq h^d \frac{p(x)}{3} \geq h^d \frac{p_{\min}}{3} \quad \textrm{for all $x \in \mc{X}$.}
-\end{equation}
-On the other hand, by setting $C_0$ to be a sufficiently large constant in~\ref{asmp:bandwidth}, we can ensure that choosing $t = n h^d p_{\min}/6$ satisfies both the inequalities~\eqref{eqn:uniform_bound_empirical_degree_1}.
+Recall the lower bound on $d_{P,h}(x)$ given in~\eqref{eqn:degree_lower_bound}.
+By setting $C_0$ to be a sufficiently large constant in~\ref{asmp:bandwidth}, we can ensure that choosing $t = n h^d p_{\min}/6$ satisfies both the inequalities~\eqref{eqn:uniform_bound_empirical_degree_1}.
 For this choice of $t$ it follows from~\eqref{eqn:degree_lower_bound} and Lemma~\ref{lem:uniform_bound_empirical_degree} that
@@
 \label{subsec:taylor_expansion}
+We begin with some notation that allows us to concisely derivatives.
+For a given $z \in \Rd$ and $s$-times differentiable function $f: \mc{X} \to \Reals$, we denote $\bigl(d_x^sf\bigr)(z) := \sum_{\abs{\alpha} = s} D^{\alpha}f(x) z^{\alpha}$.
+We also write $d^sf := \sum_{\abs{\alpha} = j} D^{\alpha}f$.
+We point out that in the first-order case $d_x^1f$ is the differential of $f$ at $x \in \mc{X}$, while $d^1f$ is the divergence of $f$.
+
 Let $u$ be a function which is $s$ times continuously differentiable at all $x \in \mc{X}$, for $k \in \mathbb{N}\setminus\{0\}$.
@@
 For notational convenience we have adopted the convention that $\sum_{j = 1}^{0} a_j = 0$.
-For a given $z \in \Rd$ and $j$-times differentiable function $f: \mc{X} \to \Reals$, we have used the notation $\bigl(d_x^jf\bigr)(z) := \sum_{\abs{\alpha} = j} D^{\alpha}f(x) z^{\alpha}$.
 Thus $\bigl(d_x^{j}f\bigr)(z)$ is a degree-$j$ polynomial---and so a $j$-homogeneous function---in $z$, meaning for any $t \in \Reals$,
@@
 \end{equation}
+In the last inequality
 

### 2021-05-26 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Updated figures for Laplacian Eigenmaps paper.

@@
 \begin{figure*}[tb]
-	\includegraphics[width=.245\textwidth]{figures/mse_by_sample_size_2d.pdf}
-	\includegraphics[width=.245\textwidth]{figures/mse_by_sample_size_3d.pdf} 
-	\includegraphics[width=.245\textwidth]{figures/mse_by_sample_size_2d_2s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/mse_by_sample_size_3d_2s.pdf} 
+	\includegraphics[width=.245\textwidth]{figures/in_sample_mse/mse_by_sample_size_1d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/in_sample_mse/mse_by_sample_size_1d_2s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/in_sample_mse/mse_by_sample_size_2d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/in_sample_mse/mse_by_sample_size_2d_2s.pdf}
 	\caption{Mean squared error of Laplacian eigenmaps (\texttt{LE}) and a spectral projection estimator (\texttt{SP}) as a function of sample size $n$.
-Each plot is on the log-log scale, and the results are averaged over 50 repetitions.
+Each plot is on the log-log scale, and the results are averaged over 400 repetitions.
 Both estimators are tuned for optimal average mean squared error.
@@
 
-In Figure~\ref{fig:fig2} we compare tests using Laplacian eigenmaps and spectral projection. \textcolor{red}{(TODO): Need to ask Siva for help.}
+We also compare the error, over a held-out test set, of Laplacian eigenmaps plus kernel smoothing to the spectral projection estimator.
+The relative performances of the two estimators, as seen in Figure~\ref{fig:fig2}, is very similar to that seen in Figure~\ref{fig:fig1}, supporting our theoretical claim that the additional error incurred by kernel smoothing of Laplacian Eigenmaps is negligible.
+\begin{figure*}[tb]
+	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_1d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_1d_2s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_2d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_2d_2s.pdf}
+	\caption{Mean squared error of Laplacian eigenmaps plus kernel smoothing (\texttt{LE+KS}) and a spectral projection estimator (\texttt{SP}) as a function of sample size $n$.
+Each plot is on the log-log scale, and the results are averaged over 200 repetitions.
+Both estimators are tuned for optimal average mean squared error.
+The black line shows the minimax rate (in slope only; the intercept is chosen to match the observed error).}
+	\label{fig:fig2}
+\end{figure*}
+
+In Figure~\ref{fig:fig3} we compare tests using Laplacian eigenmaps and spectral projection. \textcolor{red}{(TODO): Need to ask Siva for help.}
 \begin{figure*}[tb]
 	\caption{Worst-case risk Laplacian eigenmaps (\texttt{LE}) and spectral projection (\texttt{SP}) tests, as a function of sample size $n$.}
-	\label{fig:fig2}
+	\label{fig:fig3}
 \end{figure*}
@@
 	\caption{Mean squared error of Laplacian Eigenmaps as a function of hyperparameters.
-For all experiments, sample size $n = 2000$, and the results are averaged over $50$ repetitions.
+For all experiments, sample size $n = 2000$, and the results are averaged over $200$ repetitions.
 Colors show~\textcolor{red}{(?)}}
-	\label{fig:fig3}
+	\label{fig:fig4}
 \end{figure*}

### 2021-05-27 code/plot_demos/eigenmaps_tuning_parameters.R
Further update of code base for Laplacian Eigenmaps experiments.

@@ -9,10 +9,14 @@ source("plot_methods.R")
 source("misc.R")
 
 # User entered information.
-data_directory <- "data/laplacian_eigenmaps/tuning/20210525094744"
-plot_directory <- "plots/laplacian_eigenmaps/tuning" # Please change this to whichever directory you prefer.
+data_directory <- "data/laplacian_eigenmaps/out_of_sample_mse/eigenfunction_2d_2s"
+plot_directory <- "plots/laplacian_eigenmaps/tuning/misc/" # Please change this to whichever directory you prefer.
 if(!exists(plot_directory)){dir.create(plot_directory)}
-plot_n <- 1000                                       # Which sample size would you like to use?
+
+plot_n <- 3000                                       # Which sample size would you like to use?
+plot_methods <- c("laplacian_eigenmaps",
+                  "laplacian_eigenmaps_plus_kernel_smoothing",
+                  "spectral_projection")
 
 # Load data
 load(file.path(data_directory,"configs.R"))
@@ -20,35 +24,52 @@ d <- configs$d; s <- configs$s; ns <- configs$ns; methods <- configs$methods
 load(file.path(data_directory,"best_fits_by_method.R"))
 load(file.path(data_directory,"thetas.R"))
 load(file.path(data_directory,"mse.R"))
+if("test_mse.R" %in% list.files(data_directory)) load(file.path(data_directory,"test_mse.R"))
 load(file.path(data_directory,"Xs.R"))
 load(file.path(data_directory,"f0s.R"))
 load(file.path(data_directory,"Ys.R"))
 
 stopifnot(plot_n %in% ns)
-mse <- mse[plot_n %in% ns]
-methods <- methods[plot_n %in% ns]
-thetas <- thetas[plot_n %in% ns]
+mse <- mse[ns %in% plot_n][[1]][names(methods) %in% plot_methods]
+if(exists("test_mse")) test_mse <- test_mse[ns %in% plot_n][[1]][names(methods) %in% plot_methods]
+thetas <- thetas[ns %in% plot_n][[1]][names(methods) %in% plot_methods]
 
[…]

### 2021-05-27 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Updated Figures 1 and 3 of Laplacian Eigenmaps paper.

--- 2021-05-21 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
@@
 \textcolor{red}{(TODO): Include more details.}
-In this section we empirically demonstrate that Laplacian Eigenmaps is a very good alternative to spectral projection, even when $n$ is only moderately large.
+In this section we empirically demonstrate that Laplacian Eigenmaps is a reasonably good alternative to spectral projection, even when $n$ is only moderately large.
 In order to compare the two methods, in our experiments we stick to simple settings where we can compute eigenfunctions of $\Delta_P$, and thus the spectral projection estimator.
-In general, this is not so easily done: hence the appeal of Laplacian Eigenmaps.
-
-In our first experiment, we compare the mean-squared error of Laplacian eigenmaps to that of its classical counterpart.
-We vary the sample size from $n = 200$ to $n = 2000$; sample $n$ design points $X_1,\ldots,X_n$ from the uniform distribution on the cube $[-1,1]^d$; and sample responses $Y_i$ according to~\eqref{eqn:model} with regression function $f_0 = M/\sqrt{n} \cdot \sum_{k = 1}^{n/20} k^{-s/d} \psi_k$ (Note that $f_0$ changes with $n$, but always belongs to the Sobolev ball~\textcolor{red}{(?)}.) In Figure~\ref{fig:fig1} we show the mean-squared error of Laplacian eigenmaps and spectral projection as a function of $n$, for different dimensions $d$ and order of smoothness $s$.
-We see that both estimators have mean-squared error converging to zero at least as fast as the optimal rate.
-(Here the regression function $f_0$ is a little bit smoother than the worst-case Sobolev function, which explains why the minimax rate is not always saturated).
-Surprisingly, we also find that Laplacian Eigenmaps slightly but consistently outperforms its continuum counterpart.
+Of course in general, it is not easy to compute these eigenfunctions: hence the appeal of Laplacian Eigenmaps.
+
 \begin{figure*}[tb]
-	\includegraphics[width=.245\textwidth]{figures/mse_by_sample_size_2d.pdf}
-	\includegraphics[width=.245\textwidth]{figures/mse_by_sample_size_3d.pdf} 
-	\includegraphics[width=.245\textwidth]{figures/mse_by_sample_size_2d_2s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/mse_by_sample_size_3d_2s.pdf} 
-	\caption{Mean squared error of Laplacian eigenmaps (\texttt{LE}) and a spectral projection estimator (\texttt{SP}) as a function of sample size $n$.
-Each plot is on the log-log scale, and the results are averaged over 50 repetitions.
-Both estimators are tuned for optimal average mean squared error.
+	\includegraphics[width=.245\textwidth]{figures/mse/mse_by_sample_size_1d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/mse/mse_by_sample_size_1d_2s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/mse/mse_by_sample_size_2d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/mse/mse_by_sample_size_2d_2s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_1d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_1d_2s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_2d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_2d_2s.pdf}
+	\caption{Mean squared error (mse) of Laplacian eigenmaps and spectral projection estimators.
+Top row: in-sample mse of Laplacian eigenmaps (\texttt{LE}) and a spectral projection estimator (\texttt{SP}) as a function of sample size $n$.
+Bottom row: out-of-sample mse of Laplacian eigenmaps plus kernel smoothing (\texttt{LE+KS}) and a spectral projection estimator.
+Each plot is on the log-log scale, and the results are averaged over 400 repetitions.
+All estimators are tuned for optimal average mse.
 The black line shows the minimax rate (in slope only; the intercept is chosen to match the observed error).}
@@
 \end{figure*}
+
+In our first experiment, we compare the mean-squared error of Laplacian eigenmaps to that of its classical spectral projection counterpart.
+We vary the sample size from $n = 1000$ to $n = 4000$; sample $n$ design points $X_1,\ldots,X_n$ from the uniform distribution on the cube $[-1,1]^d$; and sample responses $Y_i$ according to~\eqref{eqn:model} with regression function $f_0 = M/\lambda_K^{s/2} \cdot \psi_K$ for $K \asymp n^{d/(2s + d)}$; the pre-factor $M/\lambda_K^{s/2}$ is chosen so that $|f_0|_{H^s(\mc{X})}^2 = M$.
+In Figure~\ref{fig:fig1} we show the in-sample mean-squared error of Laplacian eigenmaps and a classical spectral projection estimator as a function of $n$, for different dimensions $d$ and order of smoothness $s$.
+We see that all estimators have mean-squared error converging to zero at roughly the minimax rate.
+We also see that the mean-squared error of Laplacian Eigenmaps gets closer to that of spectral projection as $n$ gets larger.
+The fact that spectral projection outperform Laplacian Eigenmaps
+
+We also compare the error, over a held-out test set, of Laplacian eigenmaps plus kernel smoothing to the spectral projection estimator.
+The out-of-sample mean squared error of the two estimators is very similar to the in-sample mean-squared error.
+This supports our theoretical claim that the additional error incurred by kernel smoothing of Laplacian Eigenmaps is negligible.
 
@@
 In Figure~\ref{fig:fig3}, we see how the mean-squared error of Laplacian eigenmaps changes as each tuning parameter is varied.
-As suggested by our theory, properly choosing the number of eigenvectors $K$ is crucial: the mse curves have a sharply defined minimum.
-On the other hand, as a function of the graph radius parameter $\varepsilon$ the mse curves are much closer to flat.
-This is particularly true in the first-order ($s = 1$) case, and still true albeit to a lesser extent when $s = 2$.
-These findings reinforce two points we've made previously: first that Laplacian Eigenmaps should be viewed as a type of spectral projection method, in which the more important parameter to carefully tune is the number of eigenvectors $K$; second that the more regularity displayed by the regression function, the more important tuning the graph radius $\varepsilon$ becomes.
+As suggested by our theory, properly choosing the number of eigenvectors $K$ is crucial: the mean-squared error curves, as a function of $K$, always have a sharply defined minimum.
+On the other hand, as a function of the graph radius parameter $\varepsilon$ the mean-squared error curve is much closer to flat.
+This squares completely with our theory, which requires that the number of eigenvectors $K$ be much more carefully tuned that the graph radius $\varepsilon$.
+
+We also plot the out-of-sample mean squared error of Laplacian eigenmaps plus kernel smoothing, as a function of its various tuning parameters (which include the bandwidth $h$ as well as $\varepsilon$ and $K$.) Here the relationship between theory and empirics is more nuanced.
+On the one hand, empirically it seems that the optimal choice of bandwidth parameter $h$ is usually smaller than $\varepsilon$, as suggested by our theory.
+On the other hand, for Laplacian eigenmaps plus kernel smoothing we see that mean-squared error curves as a function of $K$ are often quite close to their minima even when we choose many more eigenvectors than is optimal for Laplacian eigenmaps or spectral projection.
+This is not reflected in our theory, where we require that $K$ be chosen in the same tight range as was required for Laplacian Eigenmaps to be optimal in-sample.
+However, it does make intuitive sense: extension by kernel smoothing further attenuates the noise, making the algorithm more forgiving to overfitting during the Laplacian Eigenmaps step.
 
 \begin{figure*}[tb]
-	\includegraphics[width=.245\textwidth]{figures/eigenmaps_parameters/mse_by_number_of_eigenvectors.pdf}
-	\includegraphics[width=.245\textwidth]{figures/eigenmaps_parameters/mse_by_radius.pdf} 
-	\includegraphics[width=.245\textwidth]{figures/eigenmaps_parameters_2s/mse_by_number_of_eigenvectors.pdf}
-	\includegraphics[width=.245\textwidth]{figures/eigenmaps_parameters_2s/mse_by_radius.pdf} 
-	\caption{Mean squared error of Laplacian Eigenmaps as a function of hyperparameters.
-For all experiments, sample size $n = 2000$, and the results are averaged over $50$ repetitions.
-Colors show~\textcolor{red}{(?)}}
+	\includegraphics[width=.245\textwidth]{figures/tuning/eigenfunction/mse_by_number_of_eigenvectors_1d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/tuning/eigenfunction/mse_by_radius_1d_1s.pdf} 
+	\includegraphics[width=.245\textwidth]{figures/tuning/eigenfunction/mse_by_number_of_eigenvectors_2d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/tuning/eigenfunction/mse_by_radius_2d_1s.pdf} 
+	\includegraphics[width=.245\textwidth]{figures/tuning/sobolev/mse_by_number_of_eigenvectors_1d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/tuning/sobolev/mse_by_radius_1d_1s.pdf} 
+	\caption{Mean squared error of Laplacian Eigenmaps (\textcolor{red}{red}), Laplacian Eigenmaps plus kernel smoothing (\textcolor{blue}{blue}), and spectral projection (\textcolor{green}{green}) as a function of tuning parameters.
+Top row: the same regression function $f_0$ as used in Figure~\ref{fig:fig1}.
+Bottom row: the regression function $f_0 \propto \sum_{k} 1/\lambda_k^{s/2} \psi_K$.
+For all experiments, the sample size $n = 1000$, and the results are averaged over $200$ repetitions.
+In each panel, all tuning parameters except the one being varied are set to their optimal values.
+For Laplacian Eigenmaps plus kernel smoothing, circular points and a solid line are used to denote the error as a function of the graph radius $\varepsilon$, whereas triangular points and a dashed line are used to denote the error as a function of the bandwidth $h$.}
 	\label{fig:fig3}

### 2021-06-01 code/configs/laplacian_eigenmaps/mse/eigenfunction_1d_1s.R
Added configs for Laplacian eigenmaps testing experiments.

@@ -7,7 +7,7 @@ d <- 1
 s <- 1
 M <- 2^(s - d)
 ns <- round(seq(1000,4000,length.out = 10))
-iters <- 400
+iters <- 200
 
 ### Configs for sampling. ###
 sample_X <- make_sample_uniform(d)

### 2021-06-01 code/configs/laplacian_eigenmaps/mse/eigenfunction_1d_2s.R

@@ -7,7 +7,7 @@ d <- 1
 s <- 2
 M <- 2^(s - d)
 ns <- round(seq(1000,4000,length.out = 10))
-iters <- 400
+iters <- 200
 
 ### Configs for sampling. ###
 sample_X <- make_sample_uniform(d)

### 2021-06-01 code/configs/laplacian_eigenmaps/mse/eigenfunction_2d_1s.R

@@ -7,7 +7,7 @@ d <- 2
 s <- 1
 M <- 2^(s - 2*d + 1)
 ns <- round(seq(1000,4000,length.out = 10))
-iters <- 50
+iters <- 200
 
 ### Configs for sampling. ###
 sample_X <- make_sample_uniform(d)

### 2021-06-01 code/configs/laplacian_eigenmaps/mse/eigenfunction_2d_2s.R

@@ -7,7 +7,7 @@ d <- 2
 s <- 2
 M <- 2^(s - 2*d + 1)
 ns <- round(seq(1000,4000,length.out = 10))
-iters <- 50
+iters <- 200
 
 ### Configs for sampling. ###
 sample_X <- make_sample_uniform(d)
@@ -19,12 +19,12 @@ make_f0 <- function(d,n){
 
 ### Methods. ###
 methods <- list(
-  # laplacian_eigenmaps = make_laplacian_eigenmaps,
-  spectral_projection = make_spectral_projection
-  # least_squares = make_least_squares
+  laplacian_eigenmaps = make_laplacian_eigenmaps,
+  spectral_projection = make_spectral_projection,
+  least_squares = make_least_squares
 )
 initialize_thetas <- list(
-  # initialize_laplacian_eigenmaps_thetas,
+  initialize_laplacian_eigenmaps_thetas,
+  initialize_spectral_projection_thetas,
   initialize_spectral_projection_thetas
-  # initialize_spectral_projection_thetas
 )

### 2021-06-01 code/configs/laplacian_eigenmaps/parameters/sobolev_1s_1d.R

@@ -5,7 +5,7 @@
 ### General configs.
 d <- 1
 s <- 1
-M <- 4
+M <- 4 * 2^(s - 2*d + 1)
 ns <- 1000
 iters <- 100
 

### 2021-06-01 code/configs/laplacian_eigenmaps/parameters/sobolev_1s_2d.R

@@ -5,7 +5,7 @@
 ### General configs.
 d <- 2
 s <- 1
-M <- 4
+M <- 4 * 2^(s - 2*d + 1)
 ns <- 1000
 iters <- 100
 

### 2021-06-01 code/configs/laplacian_eigenmaps/testing/eigenfunction_2d_1s.R

@@ -1,0 +1,30 @@
+#------------------------------------------------#
+# Configs for Laplacian Eigenmaps Figure 1.
+#------------------------------------------------#
+
+### General configs.
+d <- 2
+s <- 1
+M <- 2^(s - 2*d + 1) 
+ns <- round(seq(1000,4000,length.out = 10))
+iters <- 200
+B <- 100      # Number of permutations for permutation test.
+alpha <- .05   # Accepted type I error.
+
+
+### Configs for sampling. ###
+sample_X <- make_sample_uniform(d)
+make_f0s <- make_testing_eigenfunction_f0s
+
+### Methods. ###
+# Note: to speed things up, we don't take all the Ks.
+methods <- list(
+  laplacian_eigenmaps = make_laplacian_eigenmaps_test,
+  spectral_projection = make_spectral_projection_test
+  # least_squares = make_least_squares_test
+)
+initialize_thetas <- list(
+  initialize_laplacian_eigenmaps_test_thetas,
+  initialize_spectral_projection_test_thetas
+  # initialize_spectral_projection_theta_test
+)

### 2021-06-01 code/configs/laplacian_eigenmaps/testing/eigenfunction_2d_2s.R

@@ -1,0 +1,31 @@
+#------------------------------------------------#
+# Configs for Laplacian Eigenmaps Figure 1.
+#------------------------------------------------#
+
+### General configs.
+d <- 2
+s <- 2
+M <- 2^(s - 2*d + 1) 
+ns <- round(seq(1000,4000,length.out = 10))
+iters <- 200
+B <- 100      # Number of permutations for permutation test.
+alpha <- .05   # Accepted type I error.
+
+
+### Configs for sampling. ###
+sample_X <- make_sample_uniform(d)
+make_f0s <- make_testing_eigenfunction_f0s
+
+
+### Methods. ###
+# Note: to speed things up, we don't take all the Ks.
+methods <- list(
+  laplacian_eigenmaps = make_laplacian_eigenmaps_test,
+  spectral_projection = make_spectral_projection_test
+  # least_squares = make_least_squares_test
+)
+initialize_thetas <- list(
+  initialize_laplacian_eigenmaps_test_thetas,
+  initialize_spectral_projection_test_thetas
+  # initialize_spectral_projection_theta_test
+)

### 2021-06-01 code/estimation_mse_pipeline.R
Updated graph regression code base to include testing experiments.

@@ -3,19 +3,22 @@ library(dplyr)
 library(Matrix)
 library(reshape2)
 library(RSpectra)
+source("precompute_data.R")
 source("sample.R")
 source("graph.R")
-source("estimators.R")
+source("estimators.R") 
 source("manifold.R")
 source("misc.R")
+source("initialize_parameters.R")
+source("tests.R")
 
 # Please change this line to whichever config you wish to run.
-source("configs/laplacian_eigenmaps/parameters/eigenfunction_1s_2d.R")
+source("configs/laplacian_eigenmaps/mse/eigenfunction_1d_1s.R")
 
 # Options for running the pipeline
 verbose <- F
 save_fits <- F
-test_data <- T
+test_data <- F
 
 #----------------------------------------------------#
 # This is the pipeline for running regression estimation experiments.

### 2021-06-01 code/estimators.R

@@ -27,29 +27,7 @@ make_laplacian_eigenmaps <- function(theta)
     return(predict)
   }
 }
-attr(make_laplacian_eigenmaps,"precompute_data") <- function(Y,X,X_test,X_validate,theta_df){
-  rs <- unique(theta_df$r)
-  precomputed_data <- vector(mode = "list",length = length(rs))
-  names(precomputed_data) <- rs
-  for(kk in 1:length(rs)){
-    r <- rs[kk]
-    K <- thetas[[ii]][[jj]] %>% filter(r == !!r) %>% pull(K)
-    
-    # Build G_{n,r} over X.
-    G <- neighborhood_graph(X,r)
-    
-    # Get L.
-    L <- Laplacian(G)
-    
-    # Compute as many eigenvectors as we will need.
-    spectra <- get_spectra(L,max(K))
-    L_eigenvectors <- spectra$vectors[,ncol(spectra$vectors):1]
-    
-    precomputed_data[[kk]] <- L_eigenvectors
-  }
-  precomputed_data <- list(train = precomputed_data, test = NULL)
-  return(precomputed_data)
-}
+attr(make_laplacian_eigenmaps,"precompute_data") <- precompute_laplacian_eigenmaps_data
 
 make_laplacian_eigenmaps_plus_kernel_smoothing <- function(theta)
 {
@@ -221,17 +199,7 @@ make_spectral_projection <- function(theta)
     return(predict)
   }
 } 
-attr(make_spectral_projection,"precompute_data") <- function(Y,X,X_test,X_validate,theta_df){
-  precomputed_data <- vector(mode = "list",length = 3)
-  names(precomputed_data) <- c("train","validate","test")
-  
-  K <- max(theta_df$K)
-  trig_basis <- get_trigonometric_basis(d,K) # Fourier basis
-  precomputed_data[["train"]] <- if(K == 1) apply(X,1,trig_basis) else apply(X,1,trig_basis) %>% t()
-  precomputed_data[["validate"]] <- if(K == 1) apply(X_validate,1,trig_basis) else apply(X_validate,1,trig_basis) %>% t()
-  precomputed_data[["test"]] <- if(K == 1) apply(X_test,1,trig_basis) else apply(X_test,1,trig_basis) %>% t()
-  return(precomputed_data)
-}
+attr(make_spectral_projection,"precompute_data") <- precompute_least_squares_data
 
[…]

### 2021-06-01 code/initialize_parameters.R

@@ -1,0 +1,153 @@
+#------------------------------------------------------#
+# Functions to choose initial grid of tuning parameters.
+#------------------------------------------------------#
+initialize_laplacian_eigenmaps_plus_kernel_smoothing_thetas <- function(sample_X,n){
+  thetas_le <- initialize_laplacian_eigenmaps_thetas(sample_X,n)
+  rs <- unique(thetas_le$r)
+  Ks <- unique(thetas_le$K)
+  thetas <- expand.grid(r = rs, K = Ks, h = rs)
+}
+
+initialize_laplacian_eigenmaps_test_thetas <- function(sample_X,n){
+  max_K <- round(2*(M^2*n)^(2*d/(4*s + d)))
+  initialize_laplacian_eigenmaps_thetas(sample_X,n,max_K)
+}
+
+initialize_laplacian_eigenmaps_thetas <- function(sample_X,n,max_K = NULL){
+  # Choose a range of radii based on different desired minimum degree.
+  # Currently, a range so that the minimum degree is between log(n) and 2*n^{1/2}.
+  iters <- 10
+  
+  # ALDEN CHANGE
+  # OPTION MSE
+  # n_rs <- 10
+  # degrees <- round(exp(seq(log(1/4*log(n)),log(1/2*n^(1/2)),length.out = n_rs)) + 1)
+  
+  # OPTION TUNING
+  n_rs <- 10
+  degrees <- round(seq(2,60,length.out = n_rs))
+  
+  degree_dependent_rs <- matrix(ncol = n_rs, nrow = iters)
+  for(iter in 1:iters)
+  {
+    X <- sample_X(n)
+    for(jj in 1:n_rs)
+    {
+      rneighborhood <- nn2(data = X, query = X, k = degrees[jj]) 
+      degree_dependent_rs[iter,jj] <- max(rneighborhood$nn.dists[,degrees[jj]])
+    }
+    # r_connects[iter] <-  # connectivity radius
+  }
+  # Choose a range of radii intended to achieve the desired min degree.
+  rs <- colMeans(degree_dependent_rs)
+  
+  # Choose same range of eigenvectors as used for spectral projection
+  if(is.null(max_K)){
+    Ks <- initialize_spectral_projection_thetas(sample_X,n) %>% pull(K)
+  } else{
[…]

### 2021-06-01 code/misc.R

@@ -32,7 +32,7 @@ get_eigenfunction <- function(d,K){
 }
 
 get_eigenvalue <- function(d,K){
-  max_indx <- ceiling(K^(1/d))
+  max_indx <- ceiling( sqrt(d*K^(2/d)) )
   indx <- expand.grid(rep(list(0:(max_indx - 1)), d))
   eigenvalues <- apply(indx,1,FUN = function(k){sum(k^2)})
   k <- indx[order(eigenvalues)[K],,drop = FALSE]
@@ -97,6 +97,82 @@ find_best_mse <- function(mse,methods){
     best_mse[[jj]] <- data.frame(x = ns, y = best_mse_jj,sd = sd_best_mse_jj) 
   }
   return(best_mse)
+}
+
+find_critical_radius <- function(err,methods, make_f0s, type_II_error = .5, sd = T, verbose = F){
+  # Compute power
+  power <- list()
+  for(ii in 1:length(err))
+  {
+    power[[ii]] <- list()
+    for(jj in 1:length(methods))
+    {
+      power[[ii]][[jj]] <- rowMeans(err[[ii]][[jj]], dims = 2)
+    }
+  }
+  
+  # Compute critical radius
+  critical_radius <- list()
+  if(sd) iters <- dim(err[[1]][[1]])[3] # Always the same number of iterations...
+  for(jj in 1:length(methods))
+  {
+    if(sd){
+      critical_radius[[jj]] <- matrix(nrow = length(power),ncol = 3)
+      colnames(critical_radius[[jj]]) <- c("crit_radius","crit_radius_lb","crit_radius_ub")
+    } else{
+      critical_radius[[jj]] <- matrix(nrow = length(power),ncol = 1)
+      colnames(critical_radius[[jj]]) <- c("crit_radius")
+    }
+    for(ii in 1:length(power))
+    {
+      f0s <- make_f0s(d, ns[[ii]]) # First function is the zero function, so uninteresting.
+      l2_norm <- sapply(f0s,FUN = function(f){attributes(f)$l2_norm})
+      
+      # Find the easiest function for which we have at least 1 - type_II_error power
+      critical_radius_idx <- numeric(length(type_II_error))
+      critical_radius_lb_idx <- numeric(length(type_II_error))
+      critical_radius_ub_idx <- numeric(length(type_II_error))
+      for(tt in 1:length(type_II_error))
+      {
+        critical_radius_idx[tt] <- apply(power[[ii]][[jj]],2,
[…]

### 2021-06-01 code/precompute_data.R

@@ -1,0 +1,42 @@
+#----------------------------------------------------#
+# Functions to precompute data which will be used repeatedly.
+# E.g. graph eigenvectors.
+#----------------------------------------------------#
+
+precompute_least_squares_data <- function(Y,X,X_test,X_validate,theta_df){
+  precomputed_data <- vector(mode = "list",length = 3)
+  names(precomputed_data) <- c("train","validate","test")
+  
+  K <- max(theta_df$K)
+  trig_basis <- get_trigonometric_basis(d,K) # Fourier basis
+  precomputed_data[["train"]] <- if(K == 1) apply(X,1,trig_basis) else apply(X,1,trig_basis) %>% t()
+  precomputed_data[["validate"]] <- if(K == 1) apply(X_validate,1,trig_basis) else apply(X_validate,1,trig_basis) %>% t()
+  precomputed_data[["test"]] <- if(K == 1) apply(X_test,1,trig_basis) else apply(X_test,1,trig_basis) %>% t()
+  return(precomputed_data)
+}
+
+precompute_laplacian_eigenmaps_data <- function(Y,X,X_test,X_validate,theta_df){
+  if(!is.list(Y)) Y  <- list(Y)
+  rs <- unique(theta_df$r)
+  precomputed_data <- vector(mode = "list",length = length(rs))
+  names(precomputed_data) <- rs
+  for(kk in 1:length(rs)){
+    r <- rs[kk]
+    K <- thetas[[ii]][[jj]] %>% filter(r == !!r) %>% pull(K)
+    
+    # Build G_{n,r} over X.
+    G <- neighborhood_graph(X,r)
+    
+    # Get L.
+    L <- Laplacian(G)
+    
+    # Compute as many eigenvectors as we will need.
+    spectra <- get_spectra(L,max(K))
+    L_eigenvectors <- spectra$vectors[,ncol(spectra$vectors):1]
+    
+    # Store eigenvectors
+    precomputed_data[[kk]] <- L_eigenvectors 
+  }
+  precomputed_data <- list(train = precomputed_data, test = NULL)
+  return(precomputed_data)
+}

### 2021-06-01 code/sample.R

@@ -75,6 +75,78 @@ make_eigenfunction <- function(d,K,M,s){
   return(f0)
 }
 
+make_testing_eigenfunction_f0s <- function(d,n){
+  # Configuration type parameter
+  max_K <- round(2*(M^2*n)^(2*d/(4*s + d)))
+  
+  # Get eigenfunctions, and associated l2 norms
+  g0s <- lapply(1:max_K,FUN = function(k){make_eigenfunction(d,k,M,s)})
+  l2_norms <- sapply(g0s,FUN = get_l2_norm)
+  
+  # Remove anything with infinite l2 norm (constant functions), or
+  # duplicates in l2 norm.
+  g0s <- g0s[l2_norms < Inf]
+  l2_norms <- l2_norms[l2_norms < Inf]
+  g0s <- g0s[match(unique(l2_norms),l2_norms)]
+  l2_norms <- l2_norms[match(unique(l2_norms),l2_norms)]
+  
+  # Construct regression functions.
+  make_f0 <- function(g,l){
+    f0 <- function(x){2 * g(x)}
+    attr(f0,"l2_norm") <- 4 * l
+    return(f0)
+  }
+  f0s <- mapply(FUN = make_f0, g0s,l2_norms)
+  f0_null <- function(x){0}; 
+  attr(f0_null,"l2_norm") <- 0
+  f0s <- c(f0_null,f0s)
+  return(f0s)
+}
+
+make_testing_eigenfunction_mixtures_f0s <- function(d,n){
+  # Configuration type parameters
+  max_K <- round(2*(M^2*n)^(2*d/(4*s + d)))
+  n_f0s <- 50
+  
+  # Get eigenfunctions, and associated l2 norms
+  g0s <- lapply(1:max_K,FUN = function(k){make_eigenfunction(d,k,M,s)})
+  l2_norms <- sapply(g0s,FUN = get_l2_norm)
+  
+  # Remove anything with infinite l2 norm (constant functions), or
+  # duplicates in l2 norm.
+  g0s <- g0s[l2_norms < Inf]
+  l2_norms <- l2_norms[l2_norms < Inf]
+  g0s <- g0s[match(unique(l2_norms),l2_norms)]
+  l2_norms <- l2_norms[match(unique(l2_norms),l2_norms)]
+  
+  # Construct the regression functions---which are mixtures of eigenfunctions---
+  # sorted by decreasing l2 norm.
+  f0s <- list()
+  
[…]

### 2021-06-01 code/testing_error_pipeline.R

@@ -1,0 +1,147 @@
+library(RANN)
+library(dplyr)
+library(Matrix)
+library(reshape2)
+library(RSpectra)
+source("precompute_data.R")
+source("sample.R")
+source("graph.R")
+source("estimators.R") 
+source("manifold.R")
+source("misc.R")
+source("initialize_parameters.R")
+source("tests.R")
+
+# Please change this line to whichever config you wish to run.
+source("configs/laplacian_eigenmaps/testing/eigenfunction_1d_1s.R")
+
+# Options for running the pipeline
+verbose <- F
+
+#----------------------------------------------------#
+# This is the pipeline for running regression estimation experiments.
+#----------------------------------------------------#
+
+# For output.
+power <- vector(mode = "list",length = length(ns))
+err <- vector(mode = "list",length = length(ns))
+Xs <- vector(mode = "list", length = length(ns))
+f0s <- vector(mode = "list", length = length(ns))
+Ys <- vector(mode = "list", length = length(ns))
+thetas <- vector(mode = "list", length = length(ns))
+
+for(ii in 1:length(ns))
+{
+  n <- ns[ii]
+  f0s <- make_f0s(d,n)
+  power[[ii]] <- vector(mode = "list", length = length(methods))
+  err[[ii]] <- vector(mode = "list", length = length(methods))
+  
+  # Initialize tuning parameters (thetas).
+  # Note: I have granted methods the right to access the distribution P from which X is sampled
+  # in initalizing hyper-parameters.
+  thetas[[ii]] <- vector(mode = "list", length(methods))
+  
+  for(jj in 1:length(methods))
+  {
+    thetas[[ii]][[jj]] <- initialize_thetas[[jj]](sample_X, n)
+    err[[ii]][[jj]] <- array(dim = c(nrow(thetas[[ii]][[jj]]),length(f0s),iters))
+    power[[ii]][[jj]] <- matrix(nrow = nrow(thetas[[ii]][[jj]]), ncol = length(f0s))
+  }
+    
+  for(iter in 1:iters)
+  {
+    ### Data. ###
+    X <- sample_X(n)
+    Y_list <- vector(mode = "list",length = length(f0s))
+    
+    for(hh in 1:length(f0s))
+    {
+      f0 <- f0s[[hh]]
+      
+      ### Response data. ###
+      f0_evaluations <- apply(X,1,f0)
+      Y_list[[hh]] <- f0_evaluations + rnorm(n,0,1)
+    }
+    
+    ### Precomputed data. ###
[…]

### 2021-06-01 code/plot_demos/eigenmaps_tuning_parameters.R
Updated plotting methods for graph regression experiments.

@@ -9,11 +9,11 @@ source("plot_methods.R")
 source("misc.R")
 
 # User entered information.
-data_directory <- "data/laplacian_eigenmaps/out_of_sample_mse/eigenfunction_2d_2s"
-plot_directory <- "plots/laplacian_eigenmaps/tuning/misc/" # Please change this to whichever directory you prefer.
+data_directory <- "data/laplacian_eigenmaps/tuning/sobolev_1s_2d"
+plot_directory <- "plots/laplacian_eigenmaps/tuning/sobolev/" # Please change this to whichever directory you prefer.
 if(!exists(plot_directory)){dir.create(plot_directory)}
 
-plot_n <- 3000                                       # Which sample size would you like to use?
+plot_n <- 1000                                  # Which sample size would you like to use?
 plot_methods <- c("laplacian_eigenmaps",
                   "laplacian_eigenmaps_plus_kernel_smoothing",
                   "spectral_projection")
@@ -43,6 +43,7 @@ CapStr <- function(y) {
 }
 function_names <- c("eigenfunction","sobolev")
 function_name <- function_names[which(sapply(function_names,FUN = function(name){grepl(name,data_directory)}))]
+if(length(function_name) == 0) function_name = ""
 title <- paste0("d = ",d,", s = ",s,". ", CapStr(function_name),".")
 
 ## Plot 1: Mean squared error as a function of K.
@@ -79,7 +80,7 @@ ylims <- c(min(plot_mse),max(plot_mse))
 cols <- c("red","blue","green")
 stopifnot(ncol(plot_mse) <= 3)
 
-plot_name <- paste0("mse_by_number_of_eigenvectors_",d,"d_",s,"s.pdf")
+plot_name <- paste0("mse_by_number_of_eigenvectors_",plot_n,"n_",d,"d_",s,"s.pdf")
 pdf(file.path(plot_directory,plot_name))
 par(mar = c(5.1,6,4.1,4.1)) # Prevent the left hand side from getting cut off.
 # shadow plot
@@ -151,7 +152,7 @@ ylims <- c(min(plot_mse),max(plot_mse))
 cols <- c("red","blue","blue")
 stopifnot(ncol(plot_mse) <= 3)
 
-plot_name <- paste0("mse_by_radius_",d,"d_",s,"s.pdf")
+plot_name <- paste0("mse_by_radius_",plot_n,"n_",d,"d_",s,"s.pdf")
[…]

### 2021-06-01 code/plot_demos/testing.R

@@ -10,30 +10,34 @@ source("plot_methods.R")
 source("sample.R")
 
 # User entered information.
-data_directory <- "data/laplacian_eigenmaps/mse/eigenfunction_2s_2d"
-plot_directory <- "plots/laplacian_eigenmaps/mse/eigenfunction" # Please change this to whichever directory you prefer.
+data_directory <- "data/laplacian_eigenmaps/testing/20210601064335"
+plot_directory <- "plots/laplacian_eigenmaps/testing/experimental" # Please change this to whichever directory you prefer.
 if(!exists(plot_directory)){dir.create(plot_directory)}
 
+# Plotting parameters
+cols <- c("red","green")
+sd <- T
+
+# Load data
 load(file.path(data_directory,"configs.R"))
-d <- configs$d; ns <- configs$ns; s <- configs$s; M <- configs$M;
+d <- configs$d; ns <- configs$ns; s <- configs$s; M <- configs$M; make_f0s <- configs$make_f0s
 methods <- configs$methods
-load(file.path(data_directory,"best_fits_by_method.R"))
 load(file.path(data_directory,"thetas.R"))
-load(file.path(data_directory,"mse.R"))
-load(file.path(data_directory,"Xs.R"))
-load(file.path(data_directory,"f0s.R"))
-load(file.path(data_directory,"Ys.R"))
+load(file.path(data_directory,"err.R"))
+load(file.path(data_directory,"power.R"))
 
 # Subset data to methods you actually want to plot.
 plot_methods <- c("laplacian_eigenmaps",
                   "laplacian_eigenmaps_plus_kernel_smoothing",
                   "spectral_projection")
-mse <- lapply(mse,FUN = function(m){m[names(methods) %in% plot_methods]})
+err <- lapply(err,FUN = function(m){m[names(methods) %in% plot_methods]})
+power <- lapply(power,FUN = function(m){m[names(methods) %in% plot_methods]})
 methods <- methods[names(methods) %in% plot_methods]
 
-## Plot 1: Mean squared error as a function of n.
-plot_name <- paste0("mse_by_sample_size_",d,"d_",s,"s.pdf")
+# Plot critical radius.
+plot_name <- paste0("critical_radius_by_sample_size_",d,"d_",s,"s.pdf")
 pdf(file.path(plot_directory,plot_name))
 par(mar = c(5.1,6,4.1,4.1)) # Prevent the left hand side from getting cut off.
[…]

### 2021-06-01 code/plot_methods.R

@@ -96,3 +96,87 @@ plot_best_mse <- function(methods,mse,sd = T,validate = validate_mse,cols = NULL
   legend("bottomleft", legend = legend_text, col = cols, pch = 20,
          bg = "white", inset = .01, cex = 1.75)
 }
+
+# Plot the critical radius of tests.
+# Inputs:
+# -- methods: list, as in configs/testing/eigenfunction_1d_1s.R
+# -- power: list, as computed by testing_error_pipeline.R
+# -- type_II_error: the maximum tolerated level of type II error.
+# -- sd: plot standard error bars?
+# -- cols: a vector giving colors for different methods
+plot_testing_critical_radius <- function(methods,err,type_II_error = .5,
+                                         sd = T, cols = NULL, verbose = T){
+  # Calculate critical radius.
+  critical_radius <- find_critical_radius(err,methods,make_f0s,type_II_error,sd = T,
+                                          verbose)
+  
+  # Calculate slope
+  slope <- numeric(length(critical_radius))
+  names(slope) <- names(methods)
+  for(jj in 1:length(slope))
+  {
+    slope[jj] <- mse_slope(critical_radius[[jj]][,"crit_radius"],ns)
+  }
+  
+  # Calculate minimax rate
+  minimax_crit_radius <- numeric(length(ns))
+  for(ii in 1:length(ns))
+  {
+    minimax_crit_radius[ii] <- ns[ii]^(-4*s/(4*s + d))
+  }
+  minimax_crit_radius <- minimax_crit_radius * (critical_radius[[1]][1,1] / minimax_crit_radius[1])
+  
+  # Data frame for plotting
+  plot_df_list <- list()
+  for(jj in 1:length(critical_radius))
+  {
+    plot_df_list[[jj]] <- data.frame(critical_radius[[jj]])
+    name <- names(methods)[jj]
+    abbr <- case_when(
+      name == "laplacian_smoothing" ~ "LS",
+      name == "laplacian_eigenmaps" ~ "LE",
+      name == "spectral_projection" ~ "SP",
+      name == "least_squares"       ~ "LS",
+      name == "laplacian_eigenmaps_plus_kernel_smoothing" ~ "LE+KS"
+    ) 
+    names(plot_df_list)[jj] <- paste0(abbr," [Slope = ", round(slope[jj],2),"].")
+  }
[…]

### 2021-06-01 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Updated testing + tuning parameter experiments.

@@
 
-\textcolor{red}{(TODO): Include more details.}
 In this section we empirically demonstrate that Laplacian Eigenmaps is a reasonably good alternative to spectral projection, even when $n$ is only moderately large.
@@
 In our first experiment, we compare the mean-squared error of Laplacian eigenmaps to that of its classical spectral projection counterpart.
-We vary the sample size from $n = 1000$ to $n = 4000$; sample $n$ design points $X_1,\ldots,X_n$ from the uniform distribution on the cube $[-1,1]^d$; and sample responses $Y_i$ according to~\eqref{eqn:model} with regression function $f_0 = M/\lambda_K^{s/2} \cdot \psi_K$ for $K \asymp n^{d/(2s + d)}$; the pre-factor $M/\lambda_K^{s/2}$ is chosen so that $|f_0|_{H^s(\mc{X})}^2 = M$.
+We vary the sample size from $n = 1000$ to $n = 4000$; sample $n$ design points $X_1,\ldots,X_n$ from the uniform distribution on the cube $[-1,1]^d$; and sample responses $Y_i$ according to~\eqref{eqn:model} with regression function $f_0 = M/\lambda_K^{s/2} \cdot \psi_K$ for $K \asymp n^{d/(2s + d)}$; the pre-factor $M/\lambda_K^{s/2}$ is chosen so that $|f_0|_{H^s(\mc{X})}^2 = M^2$.
 In Figure~\ref{fig:fig1} we show the in-sample mean-squared error of Laplacian eigenmaps and a classical spectral projection estimator as a function of $n$, for different dimensions $d$ and order of smoothness $s$.
@@
 
-In Figure~\ref{fig:fig2} we compare tests using Laplacian eigenmaps and spectral projection. \textcolor{red}{(TODO): Need to ask Siva for help.}
+In our second experiment, we compare tests using Laplacian eigenmaps and spectral projection test statistics.
+The setup, in terms of $n$ and $P$, is the same as that of our first experiment.
+To empirically evaluate the \textcolor{red}{critical radius} $\epsilon_n(\phi,H^1(\mc{X};M))$ of a test $\phi$, we compute $\phi$ for each $f_0 \in \mc{F} \subset H^1(\mc{X};M)$, where $\mc{F}$ is a discrete subset of $H^1(\mc{X};M)$.
+For each of $b = 1,2,\ldots,100$, we compute $\epsilon_b$, the smallest value of $\epsilon$ such that $R_n(\phi,\mc{F},\epsilon_b) \geq b/100$.
+Then we take $\bar{\epsilon} = 1/100 \cdot \sum_{b = 1}^{100}\epsilon_b$ to be our empirical measure of worst-case risk.
+In Figure~\ref{fig:fig2}, we see that the critical radii of both Laplacian eigenmaps and spectral projection tests are quite close to each other, and converge to $0$ at roughly the minimax rate.
 \begin{figure*}[tb]
-	\caption{Worst-case risk Laplacian eigenmaps (\texttt{LE}) and spectral projection (\texttt{SP}) tests, as a function of sample size $n$.}
+	\includegraphics[width=.245\textwidth]{figures/testing/eigenfunction/critical_radius_by_sample_size_1d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/testing/eigenfunction/critical_radius_by_sample_size_1d_2s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/testing/eigenfunction/critical_radius_by_sample_size_2d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/testing/eigenfunction/critical_radius_by_sample_size_2d_2s.pdf}
+	\caption{Worst-case testing risk Laplacian eigenmaps (\texttt{LE}) and spectral projection (\texttt{SP}) tests, as a function of sample size $n$.
+Plots are on the same scale as Figure~\ref{fig:fig1}, and black line shows the minimax rate.
+All tests are set to have $.05$ Type I error, and are calibrated by simulation under the null.}
 	\label{fig:fig2}
@@
 	\includegraphics[width=.245\textwidth]{figures/tuning/sobolev/mse_by_radius_1d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/tuning/sobolev/mse_by_number_of_eigenvectors_2d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{figures/tuning/sobolev/mse_by_radius_2d_1s.pdf}  
 	\caption{Mean squared error of Laplacian Eigenmaps (\textcolor{red}{red}), Laplacian Eigenmaps plus kernel smoothing (\textcolor{blue}{blue}), and spectral projection (\textcolor{green}{green}) as a function of tuning parameters.
 Top row: the same regression function $f_0$ as used in Figure~\ref{fig:fig1}.
-Bottom row: the regression function $f_0 \propto \sum_{k} 1/\lambda_k^{s/2} \psi_K$.
+Bottom row: the regression function $f_0 \propto \sum_{k} 1/\lambda_k^{1/2} \psi_k$.
 For all experiments, the sample size $n = 1000$, and the results are averaged over $200$ repetitions.

### 2021-06-15 code/configs/thesis/cluster_assumption/gaussian_mixture_doppler.R
Updated code

@@ -1,21 +1,26 @@
 #------------------------------------------------#
-# Configs for Cluster Assumption plot in thesis.
+# Configs for Cluster Assumption plot in thesis, when 
+# design density is a mixture of two Gaussians.
 #------------------------------------------------#
 
 ### General configs.
 d <- 1
-ns <- round(seq(400,2000,length.out = 1))
-iters <- 1
+# ns <- round(seq(400,2000,length.out = 10))
+ns <- c(1000)
+iters <- 10
 
 ### Configs for sampling. ###
 sample_X <- function(n){
-  sep <- (log(n))^2/(2 * n)
-  make_sample_uniform_mixture(d, sep = sep)(n)
+  # Parameters for Gaussians
+  mu <- c(-1,1)
+  sigma <- c(.33,.33)
+  pi <- c(.5,.5)
+  
+  # Sample
+  sample_fxn <- make_sample_gaussian_mixture(d, mu = mu, sigma = sigma, pi = pi)
+  sample_fxn(n)
 }
-make_f0 <- function(d,n){
-  height <- 1
-  f0 <- make_stepfunction(height)
-}
+make_f0 <- make_doppler
 
 
 ### Methods. ###
@@ -24,21 +29,22 @@ methods <- list(
   laplacian_smoothing = make_laplacian_smoothing,
   least_squares = make_least_squares,
   kernel_smoothing = make_kernel_smoothing
-  )
+)
 
 ### Tuning parameters for methods. ###
 # NOTE: to minimize computation time, we choose max_K_graph smaller than max_K_least_squares
 #       This value of max_K_graph will be more than big enough for the stepfunction with cluster assumption.
-max_K_graph <- 10
+max_K_graph <- 30
 max_K_least_squares <- 30
 max_degree_kernel_smoothing <- function(n){round(log(n)^2)}
+dist_to_r <- function(dists){quantile(dists,.75)}
 initialize_thetas <- list(
   laplacian_eigenmaps = function(sample_X,n){
-    initialize_laplacian_eigenmaps_thetas(sample_X,n,max_K = max_K_graph)
-    },
+    initialize_laplacian_eigenmaps_thetas(sample_X,n,max_K = max_K_graph,dist_to_r = dist_to_r)
+  },
   laplacian_smoothing = function(sample_X,n){
-    initialize_laplacian_smoothing_thetas(sample_X,n,max_K = max_K_graph)
-    },
[…]

### 2021-06-15 code/configs/thesis/cluster_assumption/gaussian_mixture_signed_rt.R

@@ -1,20 +1,29 @@
 #------------------------------------------------#
-# Configs for Cluster Assumption plot in thesis.
+# Configs for Cluster Assumption plot in thesis, when 
+# design density is a mixture of two Gaussians.
 #------------------------------------------------#
 
 ### General configs.
 d <- 1
-ns <- round(seq(400,2000,length.out = 1))
-iters <- 1
+# ns <- round(seq(400,2000,length.out = 10))
+ns <- c(1000)
+iters <- 10
 
 ### Configs for sampling. ###
 sample_X <- function(n){
-  sep <- (log(n))^2/(2 * n)
-  make_sample_uniform_mixture(d, sep = sep)(n)
+  # Parameters for Gaussians
+  mu <- c(-1,1)
+  sigma <- c(.4,.4)
+  pi <- c(.5,.5)
+  
+  # Sample
+  sample_fxn <- make_sample_gaussian_mixture(d, mu = mu, sigma = sigma, pi = pi)
+  sample_fxn(n)
 }
 make_f0 <- function(d,n){
-  height <- 1
-  f0 <- make_stepfunction(height)
+  g <- make_sgn_rt(d,n,power = 1/4)
+  f <- function(x){4 * g(x)}
+  return(f)
 }
 
 
@@ -24,21 +33,22 @@ methods <- list(
   laplacian_smoothing = make_laplacian_smoothing,
   least_squares = make_least_squares,
   kernel_smoothing = make_kernel_smoothing
-  )
+)
 
 ### Tuning parameters for methods. ###
 # NOTE: to minimize computation time, we choose max_K_graph smaller than max_K_least_squares
 #       This value of max_K_graph will be more than big enough for the stepfunction with cluster assumption.
-max_K_graph <- 10
+max_K_graph <- 30
 max_K_least_squares <- 30
 max_degree_kernel_smoothing <- function(n){round(log(n)^2)}
+dist_to_r <- function(dists){quantile(dists,.75)}
 initialize_thetas <- list(
   laplacian_eigenmaps = function(sample_X,n){
-    initialize_laplacian_eigenmaps_thetas(sample_X,n,max_K = max_K_graph)
-    },
+    initialize_laplacian_eigenmaps_thetas(sample_X,n,max_K = max_K_graph,dist_to_r = dist_to_r)
+  },
   laplacian_smoothing = function(sample_X,n){
[…]

### 2021-06-15 code/configs/thesis/cluster_assumption/gaussian_mixture_stepfunction.R

@@ -1,16 +1,23 @@
 #------------------------------------------------#
-# Configs for Cluster Assumption plot in thesis.
+# Configs for Cluster Assumption plot in thesis, when 
+# design density is a mixture of two Gaussians.
 #------------------------------------------------#
 
 ### General configs.
 d <- 1
-ns <- round(seq(400,2000,length.out = 1))
+ns <- round(seq(400,2000,length.out = 10))
 iters <- 1
 
 ### Configs for sampling. ###
 sample_X <- function(n){
-  sep <- (log(n))^2/(2 * n)
-  make_sample_uniform_mixture(d, sep = sep)(n)
+  # Parameters for Gaussians
+  mu <- c(-1,1)
+  sigma <- c(.4,.4)
+  pi <- c(.5,.5)
+  
+  # Sample
+  sample_fxn <- make_sample_gaussian_mixture(d, mu = mu, sigma = sigma, pi = pi)
+  sample_fxn(n)
 }
 make_f0 <- function(d,n){
   height <- 1
@@ -24,7 +31,7 @@ methods <- list(
   laplacian_smoothing = make_laplacian_smoothing,
   least_squares = make_least_squares,
   kernel_smoothing = make_kernel_smoothing
-  )
+)
 
 ### Tuning parameters for methods. ###
 # NOTE: to minimize computation time, we choose max_K_graph smaller than max_K_least_squares
@@ -32,13 +39,14 @@ methods <- list(
 max_K_graph <- 10
 max_K_least_squares <- 30
 max_degree_kernel_smoothing <- function(n){round(log(n)^2)}
+dist_to_r <- function(dists){quantile(dists,.75)}
 initialize_thetas <- list(
   laplacian_eigenmaps = function(sample_X,n){
-    initialize_laplacian_eigenmaps_thetas(sample_X,n,max_K = max_K_graph)
-    },
+    initialize_laplacian_eigenmaps_thetas(sample_X,n,max_K = max_K_graph,dist_to_r = dist_to_r)
+  },
   laplacian_smoothing = function(sample_X,n){
-    initialize_laplacian_smoothing_thetas(sample_X,n,max_K = max_K_graph)
-    },
+    initialize_laplacian_smoothing_thetas(sample_X,n,max_K = max_K_graph,dist_to_r = dist_to_r)
+  },
   least_squares = function(sample_X,n){
     initialize_spectral_projection_thetas(sample_X,n,max_K = max_K_least_squares)
   },

### 2021-06-15 code/estimation_mse_pipeline.R

@@ -13,7 +13,7 @@ source("initialize_parameters.R")
 source("tests.R")
 
 # Please change this line to whichever config you wish to run.
-source("configs/thesis/cluster_assumption/step_function.R")
+source("configs/thesis/cluster_assumption/gaussian_mixture_stepfunction.R")
 
 # Options for running the pipeline
 verbose <- F
@@ -171,6 +171,8 @@ for(directory in c(save_directory))
 {
   dir.create(directory)
 }
+if(!exists("s")) s <- NULL
+if(!exists("M")) M <- NULL
 configs <- list(d = d,
                 s = s,
                 ns = ns,

### 2021-06-15 code/graph.R

@@ -1,14 +1,14 @@
-neighborhood_graph <- function(X, r, X_new = X,loop = FALSE)
+neighborhood_graph <- function(X, r, X_new = X,loop = FALSE, kernel = function(s){s <= 1})
 {
   # unpack necessary parameters
   N <- nrow(X)
   
   k_max <- round(min(4*N^(.5),N - 1))
-  A <- knn_to_neighborhood_graph(X,r,k_max,X_new,loop)
+  A <- knn_to_neighborhood_graph(X,r,k_max,X_new,loop, kernel)
   return(A)
 }
 
-knn_to_neighborhood_graph <- function(X, r, k_max, X_new = X,loop)
+knn_to_neighborhood_graph <- function(X, r, k_max, X_new = X,loop, kernel)
 {
   #--------------------#
   # Input: X (n x d matrix)
@@ -23,27 +23,38 @@ knn_to_neighborhood_graph <- function(X, r, k_max, X_new = X,loop)
   n <- nrow(X)
   m <- nrow(X_new)
   d <- ncol(X)
-  A <- Matrix(0, nrow = n, ncol = m)
-  
-  # compute distance matrix
-  rneighbors <- nn2(data = X, query = X_new, searchtype = 'radius', k = k_max, radius = r)$nn.idx
-  if( (!loop) & all(X_new == X)) rneighbors <- rneighbors[,-1]
-  rneighbors[rneighbors == 0] <- NA
-  r_neighbors_list <- as.matrix(melt(rneighbors)[,-2])
-  r_neighbors_list <- r_neighbors_list[rowSums(is.na(r_neighbors_list)) == 0,]
-  A[r_neighbors_list] <- 1
-  
-  # if k_max was not high enough, at least one vertex in A will have exactly k_max - 1 neighbors.
-  # Increase k_max and recalculate.
-  while((max(rowSums(A)) == k_max - 1) & (k_max < n - 1))
-  {
+  repeat{
     A <- Matrix(0, nrow = n, ncol = m)
-    k_max <- min(2*k_max,n - 1)
-    rneighbors <- nn2(data = X, query = X_new, searchtype = 'radius', k = k_max, radius = r)$nn.idx[,-1]
+    
+    # compute distance matrix
+    rneighborhood <- nn2(data = X, query = X_new, searchtype = 'radius', k = k_max, radius = r)
+    rneighbors <- rneighborhood$nn.idx
+    dist_to_neighbors <- rneighborhood$nn.dists
[…]

### 2021-06-15 code/initialize_parameters.R

@@ -1,5 +1,5 @@
 #------------------------------------------------------#
-# Functions to choose initial grid of tuning parameters.
+# Functions to choose initial grid of tuning parameters, for estimation.
 #------------------------------------------------------#
 initialize_laplacian_eigenmaps_plus_kernel_smoothing_thetas <- function(sample_X,n){
   thetas_le <- initialize_laplacian_eigenmaps_thetas(sample_X,n)
@@ -8,25 +8,18 @@ initialize_laplacian_eigenmaps_plus_kernel_smoothing_thetas <- function(sample_X
   thetas <- expand.grid(r = rs, K = Ks, h = rs)
 }
 
-initialize_laplacian_eigenmaps_test_thetas <- function(sample_X,n){
-  max_K <- round(2*(M^2*n)^(2*d/(4*s + d)))
-  initialize_laplacian_eigenmaps_thetas(sample_X,n,max_K)
-}
-
-initialize_laplacian_eigenmaps_thetas <- function(sample_X,n,max_K = NULL){
-  # Choose a range of radii based on different desired minimum degree.
-  # Currently, a range so that the minimum degree is between log(n) and 2*n^{1/2}.
-  iters <- 10
+initialize_laplacian_eigenmaps_thetas <- function(sample_X,n,
+                                                  n_rs = 10,
+                                                  max_degree = 60,
+                                                  max_K = max(initialize_spectral_projection_thetas$K),
+                                                  dist_to_r = max){
   
-  # ALDEN CHANGE
-  # OPTION MSE
-  # n_rs <- 10
-  # degrees <- round(exp(seq(log(1/4*log(n)),log(1/2*n^(1/2)),length.out = n_rs)) + 1)
+  degrees <- seq(2,max_degree + 1,length.out = n_rs) %>% round() %>% unique()
   
-  # OPTION TUNING
-  n_rs <- 10
-  degrees <- round(seq(2,60,length.out = n_rs))
+  # Choose parameters in a data-dependent way using Monte Carlo
+  iters <- 10 
   
+  # Choose a range of radii based on different desired degree.
   degree_dependent_rs <- matrix(ncol = n_rs, nrow = iters)
   for(iter in 1:iters)
   {
@@ -34,12 +27,11 @@ initialize_laplacian_eigenmaps_thetas <- function(sample_X,n,max_K = NULL){
[…]

### 2021-06-15 code/misc.R

@@ -93,7 +93,6 @@ find_best_mse <- function(mse,methods){
       best_mse_jj[ii] <- min(rowMeans(mse_ii_jj))
       sd_best_mse_jj[ii] <- apply(mse_ii_jj,1,sd)[which.min(rowMeans(mse_ii_jj))]/sqrt(ncol(mse_ii_jj))
     }
-    # Rescale minimax mse to match intercept with best_mse
     best_mse[[jj]] <- data.frame(x = ns, y = best_mse_jj,sd = sd_best_mse_jj) 
   }
   return(best_mse)

### 2021-06-15 code/plot_demos/thesis/cluster_assumption_mse.R

@@ -1,43 +1,0 @@
-#--------------------------------------------------#
-# Script to generate plots used for Figure 1 of Laplacian Eigenmaps paper.
-#--------------------------------------------------#
-library(dplyr)
-library(ggplot2)
-library(gridExtra)
-library(viridis)
-source("misc.R")
-source("plot_methods.R")
-source("sample.R")
-
-# User entered information.
-data_directory <- "data/thesis/cluster_assumption/stepfunction"
-plot_directory <- "plots/thesis/cluster_assumption/stepfunction" # Please change this to whichever directory you prefer.
-if(!exists(plot_directory)){dir.create(plot_directory,recursive = T)}
-
-load(file.path(data_directory,"configs.R"))
-d <- configs$d; ns <- configs$ns; s <- configs$s; M <- configs$M;
-methods <- configs$methods
-load(file.path(data_directory,"best_fits_by_method.R"))
-load(file.path(data_directory,"thetas.R"))
-load(file.path(data_directory,"mse.R"))
-load(file.path(data_directory,"Xs.R"))
-load(file.path(data_directory,"f0s.R"))
-load(file.path(data_directory,"Ys.R"))
-
-# Subset data to methods you actually want to plot.
-plot_methods <- c("laplacian_eigenmaps",
-                  "laplacian_smoothing",
-                  "laplacian_eigenmaps_plus_kernel_smoothing",
-                  "spectral_projection",
-                  "least_squares",
-                  "kernel_smoothing")
-mse <- lapply(mse,FUN = function(m){m[names(methods) %in% plot_methods]})
-methods <- methods[names(methods) %in% plot_methods]
-
-## Plot 1: Mean squared error as a function of n.
-plot_name <- paste0("mse_by_sample_size_stepfunction.pdf")
-pdf(file.path(plot_directory,plot_name))
-par(mar = c(5.1,6,4.1,4.1)) # Prevent the left hand side from getting cut off.
-plot_best_mse(methods,mse,sd = T, cols = c("red","green","blue","purple"),
-              legend = F, title = "Stepfunction.", rate = F)
-dev.off()

### 2021-06-15 code/plot_demos/thesis/cluster_assumption_plots.R

@@ -1,0 +1,86 @@
+#--------------------------------------------------#
+# Script to generate plots for cluster assumption regression fxn in thesis.
+#--------------------------------------------------#
+library(dplyr)
+library(ggplot2)
+library(gridExtra)
+library(viridis)
+source("misc.R")
+source("plot_methods.R")
+source("sample.R")
+
+# User entered information.
+data_directory <- "data/thesis/cluster_assumption/20210608233004"
+plot_directory <- "plots/thesis/cluster_assumption/gaussian_mixture_stepfunction" # Please change this to whichever directory you prefer.
+if(!exists(plot_directory)){dir.create(plot_directory,recursive = T)}
+
+# Load data
+load(file.path(data_directory,"configs.R"))
+d <- configs$d; ns <- configs$ns; s <- configs$s; M <- configs$M;
+methods <- configs$methods
+load(file.path(data_directory,"best_fits_by_method.R"))
+load(file.path(data_directory,"thetas.R"))
+load(file.path(data_directory,"mse.R"))
+load(file.path(data_directory,"Xs.R"))
+load(file.path(data_directory,"f0s.R"))
+load(file.path(data_directory,"Ys.R"))
+
+# Subset data to methods you actually want to plot.
+plot_methods <- c("laplacian_eigenmaps",
+                  "laplacian_smoothing",
+                  "least_squares",
+                  "kernel_smoothing")
+mse <- lapply(mse,FUN = function(m){m[names(methods) %in% plot_methods]})
+methods <- methods[names(methods) %in% plot_methods]
+
+## Plots 1-3: Actual and estimated regression functions.
+ii <- 1
+n <- ns[ii]
+# sep <- (log(n))^2/(2 * n) # TODO: Should have a way to grab this from data!
+# domains <- list(c(0,1/2 - sep),c(1/2 + sep,1))
+domains <- c(-1,1)
+
+# Plot 1: Regression function
+plot_name <- paste0("actual_function.pdf")
+pdf(file.path(plot_directory,plot_name))
+par(mar = c(5.1,6,4.1,4.1)) # Prevent the left hand side from getting cut off.
+plot_fxn(x = Xs[[ii]], 
+         y = Ys[[ii]],
+         f = apply(Xs[[ii]],1,f0s[[ii]]),
+         domains = domains,
+         title = "True function.")
+dev.off()
+
[…]

### 2021-06-15 code/plot_demos/thesis/cluster_assumption_tuning.R

@@ -1,0 +1,70 @@
+#--------------------------------------------------#
+# Script to generate plots analyzing impact of tuning parameters.
+#--------------------------------------------------#
+library(dplyr)
+library(ggplot2)
+library(gridExtra)
+library(viridis)
+source("plot_methods.R")
+source("misc.R")
+
+# User entered information.
+data_directory <- "data/thesis/cluster_assumption/20210607144055"
+plot_directory <- "plots/thesis/cluster_assumption/tuning/gaussian_mixture_signed_rt" # Please change this to whichever directory you prefer.
+if(!exists(plot_directory)){dir.create(plot_directory)}
+
+# Load data
+load(file.path(data_directory,"configs.R"))
+d <- configs$d; s <- configs$s; ns <- configs$ns; methods <- configs$methods
+load(file.path(data_directory,"best_fits_by_method.R"))
+load(file.path(data_directory,"thetas.R"))
+load(file.path(data_directory,"mse.R"))
+if("test_mse.R" %in% list.files(data_directory)) load(file.path(data_directory,"test_mse.R"))
+load(file.path(data_directory,"Xs.R"))
+load(file.path(data_directory,"f0s.R"))
+load(file.path(data_directory,"Ys.R"))
+
+# Subset data to methods you actually want to plot.
+plot_methods <- c("laplacian_eigenmaps",
+                  "laplacian_smoothing",
+                  "least_squares",
+                  "kernel_smoothing")
+mse <- lapply(mse,FUN = function(m){m[names(methods) %in% plot_methods]})
+thetas <- lapply(thetas,FUN = function(t){t[names(methods) %in% plot_methods]})
+
+# Plot 1: Error as a function of K
+plot_methods <- c("laplacian_eigenmaps","least_squares")
+plot_mse <- lapply(mse,FUN = function(m){m[names(methods) %in% plot_methods]})[1:9]
+plot_thetas <- lapply(thetas,FUN = function(t){t[names(methods) %in% plot_methods]})[1:9]
+
+cols <- c("red","blue")
+title <- paste0("n = ",ns,".")
+
+plot_name <- paste0("mse_by_K.pdf")
+pdf(file.path(plot_directory,plot_name))
+par(mfrow = c(2,2))
+plot_mse_by_tuning_parameter(parameter_name = "K",
+                             plot_methods,plot_thetas,plot_mse,
+                             cols,title,
[…]

### 2021-06-15 code/plot_demos/thesis/cluster_assumption_tuning_parameters.R

@@ -1,178 +1,0 @@
-#--------------------------------------------------#
-# Script to generate plots analyzing impact of LE tuning parameters.
-#--------------------------------------------------#
-library(dplyr)
-library(ggplot2)
-library(gridExtra)
-library(viridis)
-source("plot_methods.R")
-source("misc.R")
-
-# User entered information.
-data_directory <- "data/thesis/cluster_assumption/stepfunction"
-plot_directory <- "plots/thesis/cluster_assumption/tuning/stepfunction" # Please change this to whichever directory you prefer.
-if(!exists(plot_directory)){dir.create(plot_directory)}
-
-plot_n <- 1111                                 # Which sample size would you like to use?
-plot_methods <- c("laplacian_eigenmaps",
-                  "laplacian_smoothing",
-                  "laplacian_eigenmaps_plus_kernel_smoothing",
-                  "spectral_projection",
-                  "least_squares",
-                  "kernel_smoothing")
-
-# Load data
-load(file.path(data_directory,"configs.R"))
-d <- configs$d; s <- configs$s; ns <- configs$ns; methods <- configs$methods
-load(file.path(data_directory,"best_fits_by_method.R"))
-load(file.path(data_directory,"thetas.R"))
-load(file.path(data_directory,"mse.R"))
-if("test_mse.R" %in% list.files(data_directory)) load(file.path(data_directory,"test_mse.R"))
-load(file.path(data_directory,"Xs.R"))
-load(file.path(data_directory,"f0s.R"))
-load(file.path(data_directory,"Ys.R"))
-
-stopifnot(plot_n %in% ns)
-mse <- mse[ns %in% plot_n][[1]][names(methods) %in% plot_methods]
-if(exists("test_mse")) test_mse <- test_mse[ns %in% plot_n][[1]][names(methods) %in% plot_methods]
-thetas <- thetas[ns %in% plot_n][[1]][names(methods) %in% plot_methods]
-
-# Plotting parameters for all plots
-title <- "Stepfunction."
-
[…]

### 2021-06-15 code/plot_methods.R

@@ -1,12 +1,28 @@
-plot_fxn <- function(x,y,f,title = NULL,col = NULL){
-    if(is.null(col)){col <- "black"}
-    plot_df <- data.frame(x = x, y = y,f = f)
-    plot(x = plot_df$x,y = plot_df$y, 
+plot_fxn <- function(x,y,f,title = NULL,cols = "black",domains = c(0,1)){
+    # Observed values.
+    if(!is.list(domains)) domains <- list(domains)
+    domain <- c(min(unlist(domains)), max(unlist(domains)))
+    plot(x = x,y = y, 
          col = "grey34",cex = .6,
          cex.main = 2.5, cex.lab = 2.5, cex.axis = 2,
          lwd = 1.1,xlab = "",ylab = "", main = title)
-    lines(x = x[order(x)],y = f[order(x)],lwd = 1.5,
-          col = col)
+  
+    # Plot functions.
+    if(!is.list(f))
+    {
+      f <- list(f)
+    }
+    for(jj in 1:length(f))
+    {
+      for(interval in domains)
+      {
+        interval_indx <- c(x >= interval[1] & x <= interval[2])
+        plot_x <- x[interval_indx]
+        plot_f <- f[[jj]][interval_indx]
+        lines(x = plot_x[order(plot_x)], y = plot_f[order(plot_x)],lwd = 2,
+              col = cols[jj])
+      }
+    }
 }
 
 
@@ -31,7 +47,7 @@ plot_best_mse <- function(methods,mse,sd = T,validate = validate_mse,cols = NULL
     for(ii in 1:length(ns))
     {
       mse_ii_jj <- mse[[ii]][[jj]]
-      best_mse[ii] <- min(rowMeans(mse_ii_jj))
+      best_mse[ii] <- min(rowMeans(mse_ii_jj),na.rm = T)
       sd_best_mse[ii] <- apply(mse_ii_jj,1,sd)[which.min(rowMeans(mse_ii_jj))]/sqrt(ncol(mse_ii_jj))
       minimax_mse[ii] <- ns[ii]^{-2*s/(2*s + d)}
     }
@@ -182,4 +198,45 @@ plot_testing_critical_radius <- function(methods,err,type_II_error = .5,
   grid(equilogs = F, lwd = 2)
   legend("bottomleft", legend = legend_text, col = cols, pch = 20,
          bg = "white", inset = .01, cex = 1.75)
+}
+
[…]

### 2021-06-15 code/sample.R

@@ -37,12 +37,12 @@ make_sample_uniform <- function(d)
   }
 }
 
-# p(x) = 1/2 * (1{0 < x < 1/2 - sep} + 1{1/2 + sep < x < 1}).
+# Mixture of two uniforms, P = 1/2 Unif(-1,-sep) + 1/2 Unif(sep,1).
 make_sample_uniform_mixture <- function(d,sep)
 {
   stopifnot(d == 1) # For now, just 1d.
   stopifnot(sep < 1/2)
-  L <- c(0,1/2 + sep); U = c(1/2 - sep,1) # Lower and upper extremes.
+  L <- c(-1,-sep); U = c(sep,1) # Lower and upper extremes.
   sample_X <- function(n){
       class <- apply( rmultinom(n, 1, prob = c(1/2,1/2)),2,FUN = function(i){which(i != 0)})
       X <- sapply(class,FUN = function(l){runif(1,L[l],U[l])}) %>% as.matrix()
@@ -66,8 +66,14 @@ make_sample_manifold <- function(d,D)
 make_doppler <- function(d,n){
   stopifnot(d == 1)
   f0 <- function(x){
-    if(x > 0){cos(4*pi/((abs(x)^{1/3})))}
-    else{cos(pi/(abs(x)^{1/3}))}
+    cos(4*pi/abs(x)^{1/3}) 
+  }
+}
+
+make_sgn_rt <- function(d,n,power = 1/2){
+  stopifnot(d == 1)
+  f0 <- function(x){
+    sign(x)*abs(x)^(power)
   }
 }
 
@@ -195,7 +201,7 @@ make_wiggly_gaussian <- function(d,n)
 
 make_stepfunction <- function(height){
   f0 <- function(x){
-    return(height*((0 < x & x < 1/2) - (1/2 < x & x < 1)))
+    return(height*(2*(x > 0) - 1))
   }
 }
 

### 2021-06-17 notes/misc/splines_graph_laplacians_connections.tex
Part-way through editing the proof of Laplacian regularization out-of-sample.

@@
 
+To begin, we note that
+\begin{equation*}
+(S_N - S_n)f_0 = S_N(S_n^{-1} - S_N^{-1})S_nf_0 = \rho S_N\biggl(\frac{2n + 1}{(n + 1)^2}L_{n,\varepsilon} - \frac{n}{(n + 1)^2}L_{n,\varepsilon}^{(u)}\biggr)S_nf_0
+\end{equation*}
+and consequently
+\begin{equation}
+\label{pf:out_of_sample_laplacian_regularization_3}
+\|(S_N - S_n)f_0\|_{\infty} \leq \frac{\rho}{n}\|S_N\|_{\infty}\Bigl(\|L_{n,\varepsilon}S_nf_0\|_{\infty} + \|L_{n,\varepsilon}^{(u)}S_nf_0\|_{\infty}\Bigr).
+\end{equation}
+Here we recall that the $\infty$-operator norm of a matrix is $\|A\|_{\infty} = \max\{\|Au\|_{\infty}:\|u\|_{\infty} = 1\}$.
+We observe that $S_N$ and $S_n$ are contractions in $L^{\infty}(\Reals^n)$, meaning that $\|S_N\|_{\infty},\|S_n\|_{\infty} \leq 1$.
+We establish that $S_n$ is a contraction, and the exact same reasoning will hold with respect to $S_N$ as well.
+For any $u \in \Reals^n$, assuming without loss of generality that $u_1 = \|u\|_{\infty}$, we have that
+\begin{equation*}
+(S_n^{-1}u)_1 = u_1 + (\rho L_{n,\varepsilon}u)_1 \geq u_1,
+\end{equation*}
+which implies that $\|S_n^{-1}u\|_{\infty} \geq \|u\|_{\infty}$. Thus for any $u \in \Reals^n$, we also have that
+\begin{equation*}
+\|u\|_{\infty} = \|S_n^{-1}S_nu\|_{\infty} \geq \|S_nu\|_{\infty},
+\end{equation*}
+demonstrating that $S_n$ is a contraction.
+
+This takes care of the factor of $\|S_N\|_{\infty}$ in~\eqref{pf:out_of_sample_laplacian_regularization_3}.
+Now we proceed to give upper bounds on~$\|L_{n,\varepsilon}S_nf_0\|_{\infty}$ and~$\|L_{n,\varepsilon}^{(u)}S_nf_0\|_{\infty}$.
+\textcolor{red}{(TODO)}
+
 Since $S_N$ and $S_n$ are symmetric, and thus $S_NS_n = S_nS_N$, it follows that

### 2021-07-10 code/configs/thesis/mse/eigenfunction_1d_1s.R
Added code to (a) better tune hyperparameters, (b) save hat matrices for Laplacian methods, (c) have some configs for thesis.

@@ -1,0 +1,52 @@
+#------------------------------------------------#
+# Configs for Laplacian Eigenmaps Figure 1.
+#------------------------------------------------#
+
+### General configs.
+d <- 1
+s <- 1
+M <- 2^(s - d)
+ns <- round(seq(1000,2000,length.out = 10))
+iters <- 100
+
+### Configs for sampling. ###
+sample_X <- make_sample_uniform(d)
+eigenfunction_indx <- round((M^2*ns[round(length(ns)/2)])^(d/(2*s + d)))
+make_f0 <- function(d,n){
+  g0 <- make_eigenfunction(d,eigenfunction_indx,M,s)
+  f0 <- function(x){2 * g0(x)}
+}
+
+
+### Methods. ###
+methods <- list(
+  laplacian_eigenmaps = make_laplacian_eigenmaps,
+  spectral_projection = make_spectral_projection,
+  # least_squares = make_least_squares
+  laplacian_smoothing = make_laplacian_smoothing
+)
+
+initialize_le <- function(sample_X,n){
+  initialize_laplacian_eigenmaps_thetas(sample_X,n,
+                                        n_rs = 15,
+                                        max_degree = 40,
+                                        max_K = 5*eigenfunction_indx,
+                                        dist_to_r = function(x){quantile(x,.9)})}
+initialize_sp <- function(sample_X,n){
+  initialize_spectral_projection_thetas(sample_X,n,
+                                        max_K = 5*eigenfunction_indx)
+}
+initialize_ls <- function(sample_X,n){
+  initialize_laplacian_smoothing_thetas(sample_X,n,
+                                        n_rs = 10,
+                                        n_rhos = 20,
+                                        max_degree = 15,
+                                        max_K = 5*eigenfunction_indx,
+                                        dist_to_r = function(x){quantile(x,.9)},
+                                        rho_multiplier = 3)}
+initialize_thetas <- list(
+  initialize_le,
+  initialize_sp,
+  # initialize_spectral_projection_thetas
+  initialize_ls
+)

### 2021-07-10 code/configs/thesis/mse/eigenfunction_2d_1s.R

@@ -1,0 +1,52 @@
+#------------------------------------------------#
+# Configs for Laplacian Eigenmaps Figure 1.
+#------------------------------------------------#
+
+### General configs.
+d <- 2
+s <- 1
+M <- 2^(s - d)
+ns <- round(seq(1000,2000,length.out = 10))
+iters <- 100
+
+### Configs for sampling. ###
+sample_X <- make_sample_uniform(d)
+eigenfunction_indx <- round((M^2*ns[round(length(ns)/2)])^(d/(2*s + d)))
+make_f0 <- function(d,n){
+  g0 <- make_eigenfunction(d,eigenfunction_indx,M,s)
+  f0 <- function(x){2 * g0(x)}
+}
+
+
+### Methods. ###
+methods <- list(
+  laplacian_eigenmaps = make_laplacian_eigenmaps,
+  spectral_projection = make_spectral_projection,
+  # least_squares = make_least_squares
+  laplacian_smoothing = make_laplacian_smoothing
+)
+
+initialize_le <- function(sample_X,n){
+  initialize_laplacian_eigenmaps_thetas(sample_X,n,
+                                        n_rs = 15,
+                                        max_degree = 40,
+                                        max_K = 5*eigenfunction_indx,
+                                        dist_to_r = function(x){quantile(x,.9)})}
+initialize_sp <- function(sample_X,n){
+  initialize_spectral_projection_thetas(sample_X,n,
+                                        max_K = 5*eigenfunction_indx)
+}
+initialize_ls <- function(sample_X,n){
+  initialize_laplacian_smoothing_thetas(sample_X,n,
+                                        n_rs = 10,
+                                        n_rhos = 20,
+                                        max_degree = 15,
+                                        max_K = 5*eigenfunction_indx,
+                                        dist_to_r = function(x){quantile(x,.9)},
+                                        rho_multiplier = 3)}
+initialize_thetas <- list(
+  initialize_le,
+  initialize_sp,
+  # initialize_spectral_projection_thetas
+  initialize_ls
+)

### 2021-07-10 code/configs/thesis/mse/eigenfunction_3d_1s.R

@@ -1,0 +1,52 @@
+#------------------------------------------------#
+# Configs for Laplacian Eigenmaps Figure 1.
+#------------------------------------------------#
+
+### General configs.
+d <- 3
+s <- 1
+M <- 2^(s - d)
+ns <- round(seq(1000,2000,length.out = 10))
+iters <- 100
+
+### Configs for sampling. ###
+sample_X <- make_sample_uniform(d)
+eigenfunction_indx <- round((M^2*ns[round(length(ns)/2)])^(d/(2*s + d)))
+make_f0 <- function(d,n){
+  g0 <- make_eigenfunction(d,eigenfunction_indx,M,s)
+  f0 <- function(x){2 * g0(x)}
+}
+
+
+### Methods. ###
+methods <- list(
+  laplacian_eigenmaps = make_laplacian_eigenmaps,
+  spectral_projection = make_spectral_projection,
+  # least_squares = make_least_squares
+  laplacian_smoothing = make_laplacian_smoothing
+)
+
+initialize_le <- function(sample_X,n){
+  initialize_laplacian_eigenmaps_thetas(sample_X,n,
+                                        n_rs = 15,
+                                        max_degree = 40,
+                                        max_K = 5*eigenfunction_indx,
+                                        dist_to_r = function(x){quantile(x,.9)})}
+initialize_sp <- function(sample_X,n){
+  initialize_spectral_projection_thetas(sample_X,n,
+                                        max_K = 5*eigenfunction_indx)
+}
+initialize_ls <- function(sample_X,n){
+  initialize_laplacian_smoothing_thetas(sample_X,n,
+                                        n_rs = 10,
+                                        n_rhos = 20,
+                                        max_degree = 15,
+                                        max_K = 5*eigenfunction_indx,
+                                        dist_to_r = function(x){quantile(x,.9)},
+                                        rho_multiplier = 3)}
+initialize_thetas <- list(
+  initialize_le,
+  initialize_sp,
+  # initialize_spectral_projection_thetas
+  initialize_ls
+)

### 2021-07-10 code/configs/thesis/mse/eigenfunction_4d_1s.R

@@ -1,0 +1,52 @@
+#------------------------------------------------#
+# Configs for Laplacian Eigenmaps Figure 1.
+#------------------------------------------------#
+
+### General configs.
+d <- 4
+s <- 1
+M <- 2^(s - d)
+ns <- round(seq(1000,2000,length.out = 10))
+iters <- 100
+
+### Configs for sampling. ###
+sample_X <- make_sample_uniform(d)
+eigenfunction_indx <- round((M^2*ns[round(length(ns)/2)])^(d/(2*s + d)))
+make_f0 <- function(d,n){
+  g0 <- make_eigenfunction(d,eigenfunction_indx,M,s)
+  f0 <- function(x){2 * g0(x)}
+}
+
+
+### Methods. ###
+methods <- list(
+  laplacian_eigenmaps = make_laplacian_eigenmaps,
+  spectral_projection = make_spectral_projection,
+  # least_squares = make_least_squares
+  laplacian_smoothing = make_laplacian_smoothing
+)
+
+initialize_le <- function(sample_X,n){
+  initialize_laplacian_eigenmaps_thetas(sample_X,n,
+                                        n_rs = 15,
+                                        max_degree = 40,
+                                        max_K = 5*eigenfunction_indx,
+                                        dist_to_r = function(x){quantile(x,.9)})}
+initialize_sp <- function(sample_X,n){
+  initialize_spectral_projection_thetas(sample_X,n,
+                                        max_K = 5*eigenfunction_indx)
+}
+initialize_ls <- function(sample_X,n){
+  initialize_laplacian_smoothing_thetas(sample_X,n,
+                                        n_rs = 10,
+                                        n_rhos = 20,
+                                        max_degree = 15,
+                                        max_K = 5*eigenfunction_indx,
+                                        dist_to_r = function(x){quantile(x,.9)},
+                                        rho_multiplier = 3)}
+initialize_thetas <- list(
+  initialize_le,
+  initialize_sp,
+  # initialize_spectral_projection_thetas
+  initialize_ls
+)

### 2021-07-10 code/configs/thesis/mse/eigenfunction_5d_1s.R

@@ -1,0 +1,52 @@
+#------------------------------------------------#
+# Configs for Laplacian Eigenmaps Figure 1.
+#------------------------------------------------#
+
+### General configs.
+d <- 5
+s <- 1
+M <- 2^(s - d)
+ns <- round(seq(1000,2000,length.out = 10))
+iters <- 100
+
+### Configs for sampling. ###
+sample_X <- make_sample_uniform(d)
+eigenfunction_indx <- round((M^2*ns[round(length(ns)/2)])^(d/(2*s + d)))
+make_f0 <- function(d,n){
+  g0 <- make_eigenfunction(d,eigenfunction_indx,M,s)
+  f0 <- function(x){2 * g0(x)}
+}
+
+
+### Methods. ###
+methods <- list(
+  laplacian_eigenmaps = make_laplacian_eigenmaps,
+  spectral_projection = make_spectral_projection,
+  # least_squares = make_least_squares
+  laplacian_smoothing = make_laplacian_smoothing
+)
+
+initialize_le <- function(sample_X,n){
+  initialize_laplacian_eigenmaps_thetas(sample_X,n,
+                                        n_rs = 15,
+                                        max_degree = 40,
+                                        max_K = 5*eigenfunction_indx,
+                                        dist_to_r = function(x){quantile(x,.9)})}
+initialize_sp <- function(sample_X,n){
+  initialize_spectral_projection_thetas(sample_X,n,
+                                        max_K = 5*eigenfunction_indx)
+}
+initialize_ls <- function(sample_X,n){
+  initialize_laplacian_smoothing_thetas(sample_X,n,
+                                        n_rs = 10,
+                                        n_rhos = 20,
+                                        max_degree = 15,
+                                        max_K = 5*eigenfunction_indx,
+                                        dist_to_r = function(x){quantile(x,.9)},
+                                        rho_multiplier = 3)}
+initialize_thetas <- list(
+  initialize_le,
+  initialize_sp,
+  # initialize_spectral_projection_thetas
+  initialize_ls
+)

### 2021-07-10 code/estimation_mse_pipeline.R

@@ -13,11 +13,11 @@ source("initialize_parameters.R")
 source("tests.R")
 
 # Please change this line to whichever config you wish to run.
-source("configs/thesis/cluster_assumption/gaussian_mixture_stepfunction.R")
+source("configs/thesis/mse/eigenfunction_5d_1s.R")
 
 # Options for running the pipeline
-verbose <- F
-save_fits <- T
+verbose <- T
+save_fits <- F
 test_data <- F
 
 #----------------------------------------------------#

### 2021-07-10 code/estimators.R

@@ -136,8 +136,14 @@ make_laplacian_smoothing <- function(theta)
     # Get L.
     L <- Laplacian(G)
     
-    # Solve (rho*L + I)f = Y for f.
-    f_hat <- as.numeric(solve(rho * L + diag(n),Y))
+    # Solve (rho*L + I)f = Y for f,
+    # unless rho is ``too large'', in which case we set equal to mean.
+    if(rho >= 1e+12){
+      f_hat <- rep(mean(Y),length(Y))
+    }else{
+      f_hat <- as.numeric(solve(rho * L + Diagonal(n),Y))
+    }
+    
     
     # "Prediction" function
     predict <- function(X){f_hat}

### 2021-07-10 code/graph.R

@@ -97,20 +97,53 @@ Laplacian <- function(A)
 
 # A wrapper around eigs_sym, which takes care of some failure cases when the
 # matrix is poorly conditioned. 
-get_spectra <- function(A,K,sigma = 0){
+# 
+# Input: A, matrix.
+#        K, number of eigenvectors.
+#        retvec, do we want eigenvectors or just eigenvalues.
+#        tol, tolerance for error in iterative computation.
+#        sigma, for shift-and-invert
+get_spectra <- function(A,K,retvec = TRUE,tol = 1e-10,sigma = tol){
   # Compute as many eigenvectors as we will need.
   maxitr <- 10000
-  spectra <- tryCatch(eigs_sym(A,K,sigma,opts = list(maxitr = maxitr)),
+  spectra <- tryCatch(eigs_sym(A,K,sigma = sigma,
+                               which = "LM", 
+                               opts = list(maxitr = maxitr,
+                                           retvec = retvec,
+                                           tol = tol)),
                       error = function(e){
-                        test <- eigs_sym(A,K, which = "SM",opts = list(maxitr = maxitr))
+                        test <- eigs_sym(A,K, which = "SM", opts = list(maxitr = maxitr,
+                                                                        retvec = retvec,
+                                                                        tol = tol))
                         return(test)
                       })
   
+  # Sometimes, this will spit out garbage. 
+  # You can diagnose this by the presence of ridiculously large eigenvalues, 
+  # when we are looking for small eigenvalues.
+  #
+  # Check if there are any large eigenvalues, and if so, try again, decreasing
+  # the tolerance for error; this seems to work.
+  while(max(abs(spectra$values)) > 1e10)
+  {
+    tol <- tol * .001
+    sigma <- tol
+    # No need for try-catch, since we know the "LM" direction works now.
+    spectra <- eigs_sym(A,K,sigma = sigma, 
+                        which = "LM", 
+                        opts = list(maxitr = maxitr,retvec = retvec,tol = tol))
+    
+                      
+  }
+  
+  # Sometimes, this won't converge.
   # If it hasn't converged, up the number of iterations and try again.
-  while(max(abs(spectra$values)) > 2e10 | length(spectra$values) < K)
+  while(length(spectra$values) < K)
   {
     maxitr <- maxitr + 10000
-    spectra <- eigs_sym(A,K,which = "SM",opts = list(maxitr = maxitr))
+    spectra <- eigs_sym(A,K,which = "SM",opts = list(maxitr = maxitr,
[…]

### 2021-07-10 code/hat_matrices.R

@@ -1,0 +1,48 @@
+make_laplacian_smoothing_hat <- function(theta)
+{
+  r <- theta[["r"]]
+  rho <- theta[["rho"]]
+  kernel <- theta[["kernel"]]
+  laplacian_smoothing <- function(X){
+    
+    # Build G_{n,r} over X.
+    G <- neighborhood_graph(X,r)
+    
+    # Get L.
+    L <- Laplacian(G)
+    
+    # Solve (rho*L + I)^{-1},
+    # unless rho is ``too large'', in which case we set H to global mean.
+    if(rho >= 1e+12){
+      H <- matrix(rep(1/n,n^2),ncol = n,nrow = n)
+    }else{
+      H <- solve(rho * L + Diagonal(n))
+    }
+    
+    # "Prediction" function
+    return(H)
+  }
+}
+
+make_laplacian_eigenmaps_hat <- function(theta)
+{
+  r <- theta[["r"]]
+  K <- theta[["K"]]
+  kernel <- theta[["kernel"]]
+  laplacian_eigenmaps <- function(X){
+    # Build G_{n,r} over X.
+    G <- neighborhood_graph(X,r)
+    
+    # Get L.
+    L <- Laplacian(G)
+    
+    # Get spectra.
+    spectra <- get_spectra(L,K)
+    L_eigenvectors <- spectra$vectors[,ncol(spectra$vectors):1,drop = F]
+    
+    V_K <- L_eigenvectors[,1:K,drop = F] # Eigenvectors we need
+    H <- V_K %*% t(V_K)
+    
+    return(H)
+  }
+}

### 2021-07-10 code/initialize_parameters.R

@@ -10,8 +10,9 @@ initialize_laplacian_eigenmaps_plus_kernel_smoothing_thetas <- function(sample_X
 
 initialize_laplacian_eigenmaps_thetas <- function(sample_X,n,
                                                   n_rs = 10,
+                                                  n_Ks = max_K,
                                                   max_degree = 60,
-                                                  max_K = max(initialize_spectral_projection_thetas$K),
+                                                  max_K = max(initialize_spectral_projection_thetas(sample_X,n)$K),
                                                   dist_to_r = max){
   
   degrees <- seq(2,max_degree + 1,length.out = n_rs) %>% round() %>% unique()
@@ -46,13 +47,14 @@ initialize_laplacian_smoothing_thetas <- function(sample_X,n,
                                                   n_rs = 10,
                                                   n_rhos = 10,
                                                   max_degree = 60,
-                                                  max_K = max(initialize_spectral_projection_thetas$K),
-                                                  dist_to_r = max){
+                                                  max_K = max(initialize_spectral_projection_thetas(sample_X,n)$K),
+                                                  dist_to_r = max,
+                                                  rho_multiplier = 1){
   #------------------------------------------------------------#
   # Choose tuning parameters for Laplacian smoothing.
   # -- r is determined based on the minimum degree we want in the graph.
   # -- rho is chosen to be inversely proportional to the kth eigenvalue of the Laplacian,
-  #    for different values of k.
+  #    for different values of k, times a multiplier.
   #
   # Inputs:
   # -- dist_to_r: a function that goes from distances (in particular, the distance of each
@@ -63,11 +65,12 @@ initialize_laplacian_smoothing_thetas <- function(sample_X,n,
   K <- seq(1,max_K,length.out = n_rhos) %>% round() %>% unique()
   
   # Choose parameters in a data-dependent way using Monte Carlo
-  iters <- 10 
+  r_iters <- 10 
+  rho_iters <- 5
   
   # Choose a range of radii based on different desired degree.
-  degree_dependent_rs <- matrix(ncol = n_rs, nrow = iters)
-  for(iter in 1:iters)
[…]

### 2021-07-10 code/plot_demos/eigenmaps_mse.R

@@ -10,8 +10,9 @@ source("plot_methods.R")
 source("sample.R")
 
 # User entered information.
-data_directory <- "data/laplacian_eigenmaps/mse/eigenfunction_2s_2d"
-plot_directory <- "plots/laplacian_eigenmaps/mse/eigenfunction" # Please change this to whichever directory you prefer.
+data_directory <- "data/thesis/mse/eigenfunction_1s_1d"
+# plot_directory <- "plots/thesis/mse/eigenfunction/"
+plot_directory <- "../../thesis/talk/figures/mse" # Please change this to whichever directory you prefer.
 if(!exists(plot_directory)){dir.create(plot_directory)}
 
 load(file.path(data_directory,"configs.R"))
@@ -26,15 +27,18 @@ load(file.path(data_directory,"Ys.R"))
 
 # Subset data to methods you actually want to plot.
 plot_methods <- c("laplacian_eigenmaps",
-                  "laplacian_smoothing",
-                  "laplacian_eigenmaps_plus_kernel_smoothing",
-                  "spectral_projection")
+                  "laplacian_smoothing")
 mse <- lapply(mse,FUN = function(m){m[names(methods) %in% plot_methods]})
 methods <- methods[names(methods) %in% plot_methods]
 
 ## Plot 1: Mean squared error as a function of n.
 plot_name <- paste0("mse_by_sample_size_",d,"d_",s,"s.pdf")
 pdf(file.path(plot_directory,plot_name))
-par(mar = c(5.1,6,4.1,4.1)) # Prevent the left hand side from getting cut off.
-plot_best_mse(methods,mse,sd = F, cols = c("red","green","blue","purple"))
+par(mar = c(5.1,6,4.1,4.1),lwd = 3, cex.main = 4) # Prevent the left hand side from getting cut off.
+title <- paste0("d = ", d,", s = ",s,".")
+colors <-  c("purple","deepskyblue2")
+plot_methods_abbr <- c("LE","LS") # for the legend
+plot_best_mse(methods,mse,title = title, sd = F,rate = F,legend = F, cols = colors)
+legend("bottomleft", legend = plot_methods_abbr, col = colors, pch = 20, lwd = par()$lwd,
+       bg = "white", inset = .05, cex = 1.75)
 dev.off()

### 2021-07-10 code/plot_demos/eigenmaps_tuning_parameters.R

@@ -9,14 +9,12 @@ source("plot_methods.R")
 source("misc.R")
 
 # User entered information.
-data_directory <- "data/laplacian_eigenmaps/tuning/sobolev_1s_2d"
-plot_directory <- "plots/laplacian_eigenmaps/tuning/sobolev/" # Please change this to whichever directory you prefer.
+data_directory <- "data/thesis/mse/eigenfunction_1s_5d_2"
+plot_directory <- "plots/thesis/mse/eigenfunction" # Please change this to whichever directory you prefer.
 if(!exists(plot_directory)){dir.create(plot_directory)}
 
-plot_n <- 1000                                  # Which sample size would you like to use?
-plot_methods <- c("laplacian_eigenmaps",
-                  "laplacian_eigenmaps_plus_kernel_smoothing",
-                  "spectral_projection")
+plot_n <- 4000                                  # Which sample size would you like to use?
+plot_methods <- c("laplacian_eigenmaps")
 
 # Load data
 load(file.path(data_directory,"configs.R"))

### 2021-07-10 code/plot_methods.R

@@ -84,16 +84,19 @@ plot_best_mse <- function(methods,mse,sd = T,validate = validate_mse,cols = NULL
              max(plot_df_best_mse$y + plot_df_best_mse$sd,minimax_mse))
   
   # shadow plot
+  if(par()$cex.main == 1) par(cex.main = 2.5)
   plot(x = ns, xlim = xlims, ylim = ylims,
        log = "xy", xlab = "Sample size", ylab = "Mean squared error", 
-       main = title,
-       cex.main = 2.5, cex.lab = 2.5, cex.axis = 2)
+       main = title, cex.lab = 2.5, cex.axis = 2)
   
   # add points and lines
   for(jj in 1:length(methods))
   {
-    points(x = ns, y = plot_dfs_best_mse[[jj]]$y, col = cols[jj], pch = 20)
-    lines(x = ns, y = plot_dfs_best_mse[[jj]]$y, col = cols[jj],lwd = 1.5)
+    points(x = ns, y = plot_dfs_best_mse[[jj]]$y, col = cols[jj], pch = 20, cex = 2)
+    # TODO: this was hacked for backward compatibility with plots for Eigenmaps paper.
+    #       fix it once you have regenerated those plots.
+    if(par()$lwd == 1) par(lwd = 1.5)
+    lines(x = ns, y = plot_dfs_best_mse[[jj]]$y, col = cols[jj])
     if(sd)
     {
       lines(x = ns, y = plot_dfs_best_mse[[jj]]$y + plot_dfs_best_mse[[jj]]$sd, 
@@ -170,16 +173,19 @@ plot_testing_critical_radius <- function(methods,err,type_II_error = .5,
   ylims <- c(min(plot_df[,-1],minimax_crit_radius), max(plot_df[-1],minimax_crit_radius))
   
   # shadow plot
+  if(par()$cex.main == 1) par(cex.main = 2.5)
   plot(x = ns, xlim = xlims, ylim = ylims,
        log = "xy", xlab = "Sample size", ylab = "Critical radius", 
-       main = title,
-       cex.main = 2.5, cex.lab = 2.5, cex.axis = 2)
+       main = title, cex.lab = 2.5, cex.axis = 2)
   
   # add points and lines
   for(jj in 1:length(methods))
   {
     points(x = ns, y = plot_df_list[[jj]]$crit_radius, col = cols[jj], pch = 20)
[…]

### 2021-07-10 code/sample.R

@@ -13,17 +13,19 @@ make_sample_gaussian <- function(d, sigma = 1)
   }
 }
 
-make_sample_gaussian_mixture <- function(d,mu = NULL,pi = NULL,sigma = NULL)
+make_sample_gaussian_mixture <- function(d,mu = NULL,pi = NULL,sigma = NULL,truncate  = T)
 {
   if(is.null(mu)){mu <- 0}
   if(is.null(pi)){pi <- rep(1,length(mu))}
   if(is.null(sigma)){sigma <- rep(1,length(mu))}
   sample_X <- function(n){
     X <- matrix(NA,nrow = n,ncol = d)
-    while(any(is.na(X))){
-      class <- apply(rmultinom(sum(is.na(X)), 1, prob = pi),2,FUN = function(i){which(i != 0)})
-      X[which(is.na(X))] <- sapply(class,FUN = function(l){rnorm(1,mu[l],sigma[l])})
-      X[abs(X) > 1] <- NA
+    class <- apply(rmultinom(n, 1, prob = pi),2,FUN = function(i){which(i != 0)})
+    while(any(rowSums(is.na(X)) > 0)){
+      class_for_sample <- class[which(rowSums(is.na(X)) > 0)]
+      X[which(rowSums(is.na(X)) > 0),] <- 
+        sapply(class_for_sample,FUN = function(l){rnorm(d,mu[l],sigma[l])}) %>% t()
+      if(truncate){X[apply(X,1,FUN = function(x){any(abs(x) > 1)}),] <- NA}
     }
     return(X)
   }

### 2021-07-10 notes/misc/splines_graph_laplacians_connections.tex

--- 2021-05-21 notes/misc/splines_graph_laplacians_connections.tex
@@
 \end{align*}
-The first of these terms is the contribution of the noise in $Y$.
-The second of these terms is the perturbation caused by the different smoother matrices.
+The first of these terms is the contribution of the noise in the responses ${\bf Y}$.
+The second of these terms is the perturbation between smoothing with $S_N$ and smoothing with $S_n$.
 The third of these terms is the bias due to Laplacian regularization.
@@
 Now we proceed to upper bound the error induced by each of these terms.
+We will find that
+\begin{equation*}
+|\wh{a} - f_0(x_{n + 1})|^2 \leq 4\Bigl(\frac{C_1}{\delta n\varepsilon^d} + C_2\rho^2\|f_0\|_{L^{\infty}(\Omega)}^2 + \frac{C_3\rho}{\delta^2}|f_0|_{H^1(\Omega)}^2 + \frac{C_4}{\delta} \cdot \varepsilon^2 |f_0|_{H^1(\Omega)}^2\Bigr)
+\end{equation*}
+with probability at least $1 - 4\delta - C\exp(-cn\varepsilon^d)$. From here, plugging in $\varepsilon = n^{-1/(2 + d)}$ and $0 \leq \rho \leq n^{-2/(2 + d)}$ proves the claim.
 
@@
 \end{equation*}
-Also, $\|S_N\|_{\textrm{op}} \leq 1$. 
-Therefore
+It is evident that $\|S_N\|_{\textrm{op}} \leq 1$, and recalling the notation $K_{(x)} = (\eta(\|x_1 - x\|/\varepsilon),\ldots,\eta(\|x_n - x\|/\varepsilon))$ we have
 \begin{equation}
@@
 \end{equation}
-Now by Lemma~\ref{lem:max_degree_deviation}, we have that for both $K(x) = \eta(x)$ and $K = \bigl(\eta(x)\bigr)^2$, with probability at least $1 - C_2\exp\bigl(-c_2 \delta^2 \cdot n \varepsilon^d\bigr)$,
+Now by Lemma~\ref{lem:max_degree_deviation}, we have that with probability at least $1 - C_2\exp\bigl(-c_2 \cdot n \varepsilon^d\bigr)$,
 \begin{equation}
 \label{pf:out_of_sample_laplacian_regularization_1}
-\sup_{x \in \Omega} \biggl|\frac{d_{n,\varepsilon}(x; K)}{n} - d_{P,\varepsilon}(x; K)\biggr| \leq \delta \cdot \inf_{x \in \Omega} d_{P,\varepsilon}(x; K),
+\sup_{x \in \Omega} \biggl|\frac{d_{n,\varepsilon}(x)}{n} - d_{P,\varepsilon}(x; K)\biggr| \leq \frac{1}{2} \cdot \inf_{x \in \Omega} d_{P,\varepsilon}(x; K).
 \end{equation}
-and thus
+Furthermore,
 \begin{equation*}
-\sup_{x \in \Omega}\frac{1}{d_{n,\varepsilon}(x)^2} \cdot \|K_{(x)}\|_2^2 \leq \biggl[\frac{(1 + \delta)}{(1 - \delta)}\biggr]^2 \sup_{x \in \Omega} \frac{1}{nd_{P,\varepsilon}(x)^2} \int \bigl[\eta(\|x'-x\|/\varepsilon)\bigr]^2 \,dP(x') \leq \biggl[\frac{(1 + \delta)}{(1 - \delta)}\biggr]^2 \frac{\|\eta\|_{\infty}}{n p_{\min} \varepsilon^d}.
+\|K_{(x)}\|_2^2 = \sum_{i = 1}^{n} \eta\biggl(\frac{\|x_i - x_{n + 1}\|}{\varepsilon}\biggr)^2 \leq \|\eta\|_{\infty} d_{n,\varepsilon}(x_{n + 1}),
+\end{equation*}
+and combined with~\eqref{pf:out_of_sample_laplacian_regularization_1} this implies
+\begin{equation*}
+\sup_{x \in \Omega}\frac{1}{d_{n,\varepsilon}(x)^2} \cdot \|K_{(x)}\|_2^2 \leq 9 \sup_{x \in \Omega} \frac{1}{nd_{P,\varepsilon}(x)^2} \int \bigl[\eta(\|x'-x\|/\varepsilon)\bigr]^2 \,dP(x') \leq 9\frac{\|\eta\|_{\infty}}{n p_{\min} \varepsilon^d}.
 \end{equation*}
@@
 \begin{equation*}
-\Ebb_N\Bigl[\bigl(T_{n,\varepsilon}S_Nw(x_{n + 1})\bigr)^2\Bigr] \leq \biggl[\frac{(1 + \delta)}{(1 - \delta)}\biggr]^2 \frac{\|\eta\|_{\infty}}{n p_{\min} \varepsilon^d}.
+\Ebb_N\Bigl[\bigl(T_{n,\varepsilon}S_Nw(x_{n + 1})\bigr)^2\Bigr] \leq 9\frac{\|\eta\|_{\infty}}{n p_{\min} \varepsilon^d}.
 \end{equation*}
-Thus by Markov's inequality, with probability at least $1 - \beta - C_2\exp\bigl(-c_2 \delta^2 \cdot n \varepsilon^d\bigr)$
+Thus by Markov's inequality, with probability at least $1 - \delta - C\exp\bigl(-c n \varepsilon^d\bigr)$
 \begin{equation*}
-\bigl(T_{n,\varepsilon}S_Nw(x_{n + 1})\bigr)^2 \leq \frac{1}{\beta}\biggl[\frac{(1 + \delta)}{(1 - \delta)}\biggr]^2 \frac{\|\eta\|_{\infty}}{n p_{\min} \varepsilon^d} = C_1\frac{1}{n\varepsilon^d}.
+\bigl(T_{n,\varepsilon}S_Nw(x_{n + 1})\bigr)^2 \leq \frac{9}{\delta}\frac{\|\eta\|_{\infty}}{n p_{\min} \varepsilon^d} = C_1\frac{1}{\delta n\varepsilon^d}.
 \end{equation*}
@@
 \begin{equation*}
-|T_{n,\varepsilon}u(x_{n + 1})|^2 = \biggl|\frac{1}{d_{n,\varepsilon}(x_{n + 1})} \sum_{i = 1}^{n} u_i \eta\biggl(\frac{\|x_i - x_{n + 1}\|}{\varepsilon}\biggr)\biggr|^2 \leq \frac{1}{d_{n,\varepsilon}(x_{n + 1})} \sum_{i = 1}^{n} |u_i|^2 \eta\biggl(\frac{\|x_{i} - x_{n + 1}\|}{\varepsilon}\biggr) \leq \|u\|_{\infty}^2.
+|T_{n,\varepsilon}u(x_{n + 1})|^2 = \biggl|\frac{1}{d_{n,\varepsilon}(x_{n + 1})} \sum_{i = 1}^{n} u_i \eta\biggl(\frac{\|x_i - x_{n + 1}\|}{\varepsilon}\biggr)\biggr|^2 \leq \frac{1}{d_{n,\varepsilon}(x_{n + 1})} \sum_{i = 1}^{n} |u_i|^2 \eta\biggl(\frac{\|x_{i} - x_{n + 1}\|}{\varepsilon}\biggr) \leq \|u\|_{\infty}^2,
 \end{equation*}
+where in the last inequality we write $\|u\|_{\infty} := \max_{i = 1,\ldots,n} |u_i|$ for the $L^{\infty}(\Reals^n)$ norm.
 Thus to upper bound Term 2 we focus our attention on upper bounding $\|(S_N - S_n)f_0\|_{\infty}$.
 
-Since $S_N$ and $S_n$ are symmetric, and thus $S_NS_n = S_nS_N$, it follows that
+To begin with, we note that
 \begin{equation*}
-(S_N - S_n)f_0 = (S_NS_nS_n^{-1} - S_nS_NS_N^{-1})f_0 = S_N S_n(S_n^{-1} - S_N^{-1})f_0.
+(S_N - S_n)f_0 = S_N(S_n^{-1} - S_N^{-1})S_nf_0 = \rho S_N\biggl(\frac{2n + 1}{(n + 1)^2}L_{n,\varepsilon} - \frac{n}{(n + 1)^2}L_{n,\varepsilon}^{(u)}\biggr)S_nf_0 
 \end{equation*}
-Consequently we can upper bound the $\infty$-norm $\|(S_N - S_n)f_0\|_{\infty}$ by
-\begin{equation*}
-\|(S_N - S_n)f_0\|_{\infty} \leq \|S_N\|_{\infty} \|S_n\|_{\infty} \|(S_n^{-1} - S_N^{-1})f_0\|_{\infty}.
-\end{equation*}
-Here we recall that the $\infty$-operator norm of a matrix is $\|A\|_{\infty} = \max\{\|Au\|_{\infty}:\|u\|_{\infty} = 1\}$.
+and consequently
+\begin{align}
+\label{pf:out_of_sample_laplacian_regularization_3}
+\|(S_N - S_n)f_0\|_{\infty} & \leq \frac{C\rho}{n}\|S_N\|_{\infty}\Bigl(\|L_{n,\varepsilon}S_nf_0\|_{\infty} + \|L_{n,\varepsilon}^{(u)}S_nf_0\|_{\infty}\Bigr) \nonumber \\
+& \leq \frac{C\rho}{n}\|S_N\|_{\infty}\Bigl(\|L_{n,\varepsilon}\|_{\infty} + \|L_{n,\varepsilon}^{(u)}\|_{\infty}\Bigr) \|S_n\|_{\infty} \|f_0\|_{\infty} \nonumber \\
+& \leq \frac{C\|f_0\|_{L^{\infty}(\Omega)}\rho}{n}\|S_N\|_{\infty}\Bigl(\|L_{n,\varepsilon}\|_{\infty} + \|L_{n,\varepsilon}^{(u)}\|_{\infty}\Bigr) \|S_n\|_{\infty} 
+\end{align}
+Here we have used that $\|f\|_{\infty} \leq \|f\|_{L^{\infty}(\mc{X})} \leq M$, and we recall that the $\infty$-operator norm of a matrix is $\|A\|_{\infty} = \max\{\|Au\|_{\infty}:\|u\|_{\infty} = 1\}$.
+It remains to give estimates on each of $\|S_N\|_{\infty}, \|S_n\|_{\infty}, \|L_{n,\varepsilon}\|_{\infty}$ and $\|L_{n,\varepsilon}^{(u)}\|_{\infty}$.
 
-We first upper bound $\|(S_n^{-1} - S_N^{-1})f_0\|_{\infty}$, using the decomposition
-\begin{equation*}
-S_n^{-1} - S_N^{-1} = \rho \biggl[\biggl(1 - \frac{n^2}{N^2}\biggr)L_{n,\varepsilon} - \frac{n}{N^2}L_{n,\varepsilon}^{(u)}\biggr],
-\end{equation*}
-and therefore
-\begin{equation}
-\label{pf:out_of_sample_laplacian_regularization_2}
-\|(S_n^{-1} - S_N^{-1})f_0\|_{\infty} \leq \frac{\rho}{n} \Bigl(\|L_{n,\varepsilon}f_0\|_{\infty} + \|L_{n,\varepsilon}^{(u)}f_0\|_{\infty} \Bigr).
-\end{equation}
-From H\"{o}lder's inequality and the Lipschitz property of $f_0$, we deduce both that
-\begin{align*}
-\|L_{n,\varepsilon}f_0\|_{\infty} & = \max_{i = 1,\ldots,n} \biggl| \frac{1}{n\varepsilon^{d + 2}}\sum_{j = 1}^{n} \bigl(f_0(x_i) - f_0(x_j)\bigr) \eta\biggl(\frac{\|x_i - x_j\|}{\varepsilon}\biggr) \biggr| \\
-& \leq \max_{i = 1,\ldots,n}  \frac{M}{n\varepsilon^{d + 1}}\sum_{j = 1}^{n} \eta\biggl(\frac{\|x_i - x_j\|}{\varepsilon}\biggr) \\
-& \leq M\frac{1}{\varepsilon^{d + 1}},
-\end{align*}
-and
-\begin{align*}
-\|L_{n,\varepsilon}^{(u)}f_0\|_{\infty} & = \max_{i = 1,\ldots,n} \biggl| \frac{1}{\varepsilon^{d + 2}} \sum_{j = 1}^{n} \bigl(f_0(x_i) - f_0(x_j)\bigr) W_{n,ij}^{(u)}\biggr| \\
-& \leq \max_{i = 1,\ldots,n} \frac{M}{\varepsilon^{d + 1}} \sum_{j = 1}^{n} |W_{n,ij}^{(u)}|\\
-& \leq \frac{M}{\varepsilon^{d + 1}} \|\eta\|_{\infty},
-\end{align*}
-where we recall the definition of $W_{n}^{(u)}$ from~\eqref{eqn:induced_insample_graph}. Plugging these upper bounds back into~\eqref{pf:out_of_sample_laplacian_regularization_2} gives
-\begin{equation*}
-\|(S_n^{-1} - S_N^{-1})f_0\|_{\infty} \leq 2 \|\eta\|_{\infty}\frac{\rho}{n\varepsilon^{d + 1}}M.
-\end{equation*}
-On the other hand, it is not hard to see that $S_N$ and $S_n$ are contractions in $L^{\infty}(\Reals^n)$, meaning that $\|S_N\|_{\infty},\|S_n\|_{\infty} \leq 1$.
+First, we observe that $S_N$ and $S_n$ are contractions in $L^{\infty}(\Reals^n)$, meaning that $\|S_N\|_{\infty},\|S_n\|_{\infty} \leq 1$.
 We establish that $S_n$ is a contraction, and the exact same reasoning will hold with respect to $S_N$ as well.
@@
 
-Plugging these upper bounds back into~\eqref{pf:out_of_sample_laplacian_regularization_2}, we conclude that
+On the other hand, for any $v \in \Reals^n$ such that $\|v\|_{\infty} = 1$,
+\begin{align*}
+\|L_{n,\varepsilon}f_0\|_{\infty} & = \max_{i = 1,\ldots,n} \biggl| \frac{1}{n\varepsilon^{d + 2}}\sum_{j = 1}^{n} \bigl(v_i - v_j\bigr) \eta\biggl(\frac{\|x_i - x_j\|}{\varepsilon}\biggr) \biggr| \\
+& \leq \max_{i = 1,\ldots,n}  \frac{2}{n\varepsilon^{d + 2}}\sum_{j = 1}^{n} \eta\biggl(\frac{\|x_i - x_j\|}{\varepsilon}\biggr) \\
+& \leq \frac{2}{\varepsilon^{d + 2}},
+\end{align*}
+and
+\begin{align*}
+\|L_{n,\varepsilon}^{(u)}f_0\|_{\infty} & = \max_{i = 1,\ldots,n} \biggl| \frac{1}{\varepsilon^{d + 2}} \sum_{j = 1}^{n} \bigl(v - v\bigr) W_{n,ij}^{(u)}\biggr| \\
+& \leq \max_{i = 1,\ldots,n} \frac{2}{\varepsilon^{d + 1}} \sum_{j = 1}^{n} |W_{n,ij}^{(u)}|\\
+& \leq \frac{2}{\varepsilon^{d + 2}} \|\eta\|_{\infty},
+\end{align*}
+so that $\|L_{n,\varepsilon}f_0\|_{\infty}, \|L_{n,\varepsilon}^{(u)}f_0\|_{\infty} \leq C\varepsilon^{-(2 + d)}$.
+Plugging these estimates back into~\eqref{pf:out_of_sample_laplacian_regularization_3}, we conclude that
 \begin{equation*}
-\bigl(T_{n,\varepsilon}(S_N - S_n)f_0(x_{n + 1})\bigr)^2 \leq \|(S_N - S_n)f_0\|_{\infty}^2 \leq 4 \|\eta\|_{\infty}^2\biggl(\frac{\rho}{n\varepsilon^{d + 1}}\biggr)^2M^2 =: C_2 \biggl(\frac{\rho}{n\varepsilon^{d + 1}}\biggr)^2M^2.
+\bigl(T_{n,\varepsilon}(S_N - S_n)f_0(x_{n + 1})\bigr)^2 \leq \Bigl(\|(S_N - S_n)f_0\|_{\infty}\Bigr)^2 \leq \biggl(\frac{C\rho\|f_0\|_{L^{\infty}(\Omega)}}{n\varepsilon^{d + 2}}\biggr)^2 = C_2\rho^2\|f_0\|_{L^{\infty}(\Omega)}^2.
 \end{equation*}
@@
 \end{equation*}
-We have show in \textcolor{red}{our paper on Laplacian eigenmaps (equation 106)} that for any $u \in \Reals^n$,
+We have shown in \textcolor{red}{our paper on Laplacian eigenmaps (equation 106)} that for any $u \in \Reals^n$,
 \begin{equation*}
@@
 \end{equation*}
-and that under the assumption~\ref{asmp:design}, for any Sobolev function $f \in H^1(\Omega)$, with probability at least $1 - \delta$
+and that under Assumption~\ref{asmp:design}, for any Sobolev function $f \in H^1(\Omega)$, with probability at least $1 - \delta$
 \begin{equation*}
@@
 \end{equation*}
-Combining these two results, we have that
+Combining these results, we have that
 \begin{equation*}
@@
 \begin{equation*}
-\|T_{n,\varepsilon}f_0 - f_0\|_P^2 \leq \frac{C}{\delta} \cdot \biggl(\frac{\varepsilon^2}{n\varepsilon^d} + \varepsilon^2\biggr) |f|_{H^1(\Omega)}^2 \leq \frac{C_4}{\delta} \cdot \varepsilon^2 |f|_{H^1(\Omega)}^2
+\Ebb_n\Bigl[\Bigl(T_{n,\varepsilon}f_0(x_{n + 1}) - f_0(x_{n + 1})\Bigr)^2\Bigr] = \|T_{n,\varepsilon}f_0 - f_0\|_P^2 \leq \frac{C}{\delta} \cdot \biggl(\frac{\varepsilon^2}{n\varepsilon^d} + \varepsilon^2\biggr) |f|_{H^1(\Omega)}^2 \leq \frac{C_4}{\delta} \cdot \varepsilon^2 |f|_{H^1(\Omega)}^2
 \end{equation*}

### 2021-08-13 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Fixed all of Siva's small changes. Comprehensive changes made to introduction of Laplacian Eigenmaps paper.

@@
 
-\section{Introduction}
-\label{sec:introduction}
-
-Suppose we observe data $X_1,\ldots,X_n \in \Reals^d$, sampled independently from an unknown distribution $P$.
-As a replacement for $P$, we form a geometric graph $G$ over the observed data, with vertices at $X_1,\ldots,X_n$ and weighted edges $W_{ij}$ corresponding to proximity between samples $X_i$ and $X_j$.
-Geometric graphs encode both local information and global information about $P$ in an extremely general manner.
-For this reason they have been leveraged to conduct many different fundamental statistical tasks, such as clustering, manifold learning, semi-supervised learning, classification, and regression \textcolor{red}{(TODO): references}.
-Substantial theoretical progress has been made regarding the consistency of \emph{graph-based learning}, that is, learning algorithms defined with respect to geometric graphs \textcolor{red}{(TODO): references}.
-This work sheds light on why such procedures work, by showing that they converge to interesting \emph{continuum} limits as $n \to \infty$.
-However, thus far little has been said regarding the optimality of graph-based learning methods, even for classic statistical tasks.
-
-In this paper we focus on the theoretical statistical properties of a particularly popular graph-based learning method for regression.
-We assume that in addition to the design points $X_1,\ldots,X_n$ one observes real-valued responses $Y_1,\ldots,Y_n$, and seeks to learn the unknown regression function $f_0(x) := E[Y|X = x]$.
-The specific graph-based method we study is \emph{Laplacian eigenmaps}, first introduced by \cite{belkin03a}, which projects the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto the span of eigenvectors of a graph Laplacian.
-We focus on the the unnormalized graph Laplacian $L$, which is a difference operator acting on vectors $u \in \Reals^n$ as follows,
-\begin{equation}
-\label{eqn:graph_laplacian}
-(Lf)_i = \sum_{j = 1}^{n} (u_i - u_j)W_{ij}. 
-\end{equation}
-The graph Laplacian $L$ is a discrete approximation to a weighted continuum Laplacian operator $\Delta_P$, defined when $P$ admits a differentiable density $p$ as
-\begin{equation}
-\label{eqn:fokker_planck_1}
-\Delta_Pf= -\frac{1}{p} \mathrm{div}(p^2 \nabla f).
-\end{equation}
-The eigenvectors of $L$ form an orthonormal basis of $\Reals^n$, and serve as estimates of eigenfunctions of $\Delta_P$.
-Their corresponding eigenvalues are likewise estimates of eigenvalues of $\Delta_P$, and give a notion of smoothness to each eigenvector---roughly speaking, the smaller the eigenvalue, the smoother the corresponding eigenvector.
-
-We can therefore view Laplacian eigenmaps as a twist on a very classical approach to nonparametric regression: \emph{spectral projection}, or more generally \emph{orthogonal series}, regression.
-Classically, in orthogonal series regression one fixes a reference measure $Q$, takes an ordered orthonormal basis $\psi_1,\psi_2,\ldots$ of $L^2(Q)$, computes empirical Fourier coefficients $\dotp{{\bf Y}}{\psi_k}_n$, and uses the first few terms in the resulting Fourier series to construct an estimator.
-Regression by spectral projection is a special case of this general setup, where one takes $\psi_k$ to be the $k$th eigenfunction of the continuum Laplacian operator $\Delta_Q$.
-In contrast to this classical approach---in which the reference measure $Q$ and resulting Laplacian $\Delta_Q$ are determined a priori---in Laplacian eigenmaps the eigenvectors of the graph Laplacian serve as the basis.
-These eigenvectors are data-dependent objects, and adapt to the geometry of the unknown design distribution $P$ in a rich manner.
-For instance, they respect the cluster structure of $P$, meaning that if the density $p$ has multiple connected components, the first few eigenvectors of $L$ will (with high probability) be piecewise constant over each such component \textcolor{blue}{(vonLuxburg 2009)}.
-On the other hand if $P$ is supported on some low-dimensional manifold, the graph Laplacian eigenvectors concentrate around the eigenfunctions of a manifold Laplace-Beltrami operator, and thus give a principled embedding of potentially high dimensional design points $X_i$ into a lower dimensional space \citep{belkin03a}.
-
-Thus Laplacian eigenmaps is a data-dependent alternative to classical spectral projection estimators.
-This data-dependency is appealing.
-Classical spectral projection estimators possess attractive theoretical properties for nonparametric regression---in particular, they are minimax rate-optimal for regression over H\"{o}lder and Sobolev spaces \cite{tsybakov08,johnstone2011,gine16}---but in practice suffer from some serious drawbacks.
-A basic difficulty is that finding the eigenfunctions of $\Delta_Q$ is in general non-trivial, so that it may not be possible to compute the estimator in the first place.
-Moreover, spectral projection estimators make sense (and are minimax optimal) only if the reference measure $Q$ is very close or equal to $P$, since otherwise the basis functions are orthogonalized with respect to the wrong measure.
-For these reasons, such estimators are typically proposed and studied under very restrictive conditions on the design points: for instance, that they are equally spaced fixed grid points, or that they are random but uniformly distributed on the unit cube $[0,1]^d$.
-There do exist fixes to the issues just raised, for instance using nonparametric least-squares.
-We will compare Laplacian eigenmaps to nonparametric least squares in more detail later, and for now merely point out that it fundamentally changes the estimator under consideration.
-\textcolor{red}{(For Ryan + Siva): I attempted to edit this but still don't like it.}
-
-In contrast, Laplacian eigenmaps directly approximates a spectral projection method: it projects the responses onto an orthogonal basis (eigenvectors of the graph Laplacian) that approximates the smooth eigenfunctions of $\Delta_P$.
-Laplacian eigenmaps is perfectly well-defined, and indeed straightforward to compute, when the design $P$ is unknown.
-This includes the situation where $P$ may be non-uniform, or even concentrated on a low-dimensional manifold.
-On the other hand, because graph Laplacian eigenvectors depend on the random design points $X_1,\ldots,X_n$ in \textcolor{red}{(non-trivial ways)}, it is substantially more difficult to analyze Laplacian eigenmaps than to analyze classical spectral projection methods.
-For this reason, the theoretical statistical properties of Laplacian eigenmaps, and in particular its optimality as a method for nonparametric regression with random design, remain poorly understood.
-
-\paragraph{Our contributions.} The primary contribution of our paper is to fill this theoretical gap, by answering the following question:
-
-\begin{quote}
-	Is Laplacian eigenmaps an optimal method for nonparametric regression over Sobolev functions?
-\end{quote}
-
-Broadly seeking, our answer is yes.
-We consider two different data models, one of which assumes the support $\mc{X}$ of the distribution $P$ is a full-dimensional, flat Euclidean domain, and the other of which assumes that $\mc{X}$ is a manifold of small intrinsic dimension $m$.
-We show that when the regression function $f_0$ is smooth, in the Sobolev sense of having weak derivatives bounded in $\Leb^2(\mc{X})$ norm, Laplacian eigenmaps methods are statistically minimax optimal, for both estimation and goodness-of-fit testing.
-Our statements hold for different relations between the dimension $d$ (or $m$) and number of derivatives $s$, depending on the problem (estimation or testing), and are summarized in Tables~\ref{tbl:estimation_rates} and~\ref{tbl:testing_rates}.
-\begin{table}
-	\begin{center}
-		\begin{tabular}{p{.2\textwidth} | p{.14\textwidth} p{.12\textwidth} }
-			Smoothness order & Flat Euclidean (Model~\ref{def:model_flat_euclidean}) & Manifold (Model~\ref{def:model_manifold}) \\
-			\hline
-			$s \leq 3$ & $\bm{n^{-2s/(2s + d)}}$ & $\bm{n^{-2s/(2s + m)}}$ \\
-			$s > 3$  & $\bm{n^{-2s/(2s + d)}}$ & $n^{-6/(6 + m)}$
-		\end{tabular}
-	\end{center}
-	\caption{Summary of estimation rates over Sobolev balls.
-Bold font marks minimax optimal rates.
-In each case, rates hold for all $d \in \mathbb{N}$ (under Model~\ref{def:model_flat_euclidean}), and for all $m \in \mathbb{N}, 1 < m < d$ (under Model~\ref{def:model_manifold}).
-Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps estimator is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $n^{-1/2} \lesssim M \lesssim n^{1/d}$.}
-	\label{tbl:estimation_rates}
-\end{table}
-
-\begin{table}
-	\begin{center}
-		\begin{tabular}{p{.175\textwidth} p{.175\textwidth} | p{.14\textwidth} p{.12\textwidth} }
-			Smoothness order & Dimension & Flat Euclidean (Model~\ref{def:model_flat_euclidean}) & Manifold (Model~\ref{def:model_manifold}) \\
-			\hline
-			\multirow{2}{*}{$s = 1$} & $\dim(\mc{X}) < 4$ & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
-			& $\dim(\mc{X}) \geq 4$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
-			\hline
-			\multirow{3}{*}{$s = 2$ or $3$} & $\dim(\mc{X}) \leq 4$  & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
-			& $4 <\dim(\mc{X}) < 4s$  & $n^{-2s/(2(s - 1) + d)}$ & $n^{-2s/(2(s - 1) + m)}$\\
-			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
-			\hline
-			\multirow{3}{*}{$s > 3$} & $\dim(\mc{X}) \leq 4$ & $\bm{n^{-4s/(4s + d)}}$ & $n^{-12/(12 + d)}$ \\
-			& $4 < \dim(\mc{X}) < 4s$ & $n^{-2s/(2(s - 1) + d)}$ & $n^{-6/(4 + m)}$ \\
-			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
-		\end{tabular}
-	\end{center}
-	\caption{Summary of Laplacian eigenmaps testing rates over Sobolev balls.
-Bold font marks minimax optimal rates.
-Rates when $d > 4s$ assume that $f_0 \in L^4(P,M)$.
-Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $n^{-1/2} \lesssim M \lesssim n^{1/d}$.}
-	\label{tbl:testing_rates}
-\end{table}
-
-\paragraph{Related Work.}
-There is an incredible amount of work on the properties of classical spectral projection methods, which go far beyond their optimality for nonparametric regression.
-We will not attempt to summarize this work, and instead refer to \citet{wasserman2006,gyorfi2006,tsybakov08,johnstone2011,gine16} and the references therein.
-
-More recent work has considered regression on a fixed graph, where one treats the design points $x_1,\ldots,x_n$ as vertices in a fixed graph $G$, and carries out inference with respect to the function evaluations $(f_0(x_i))_{i = 1}^{n}$ By this point there exists a relatively mature theory describing this setting.
-Tight upper bounds have been established that certify the optimality of graph-based methods for both nonparametric estimation \citep{wang2016,hutter2016,sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}) and testing (e.g., \citet{sharpnack2010identifying,sharpnack2013b,sharpnack2013,sharpnack2015} over different ``function'' classes (in quotes because these classes really model the $n$-dimensional vector of evaluations) We call particular attention to~\citet{sadhanala16} and~\citet{sharpnack2015}, who analyze the Laplacian eigenmaps estimator and test statistic, respectively.
-This setting is quite general, because the graph need not be a geometric graph defined on a vertex set which belongs to Euclidean space.
-On the other hand, in many situations it may be somewhat to unnatural to assume that the design points are a priori fixed, and that the regression function $f_0$ exhibits ``smoothness'' over this fixed design.
-Instead, it may be more reasonable to adopt the \emph{random design} perspective that we work in, and assume that the regression function $f_0$ exhibits a more classical notion of smoothness.
-
-However, as already mentioned, when the design points are random so too are the graph Laplacian eigenvectors, and grasping their properties is in general non-trivial.
-For this reason, there has not been much analysis of random design nonparametric regression using Laplacian eigenmaps.
-\cite{zhou2011} consider the Laplacian eigenmaps estimator, but in the semi-supervised setting, where one additionally observes unlabeled design points $X_{n + 1},\ldots,X_{n + m}$.
-Their analysis assumed that for a fixed number of labeled samples $n$, the number of unlabeled samples $m$ grows to infinity.
-In this case the eigenvectors of the graph Laplacian converge to eigenfunctions of a continuum Laplacian, and the analysis of the resulting estimator is identical to that of a classical spectral projection estimator.
-\cite{lee2016} consider the \emph{diffusion maps} estimator---which uses the eigenvectors of a different normalization of the graph Laplacian $L$---in both the supervised and semi-supervised setups.
-In the supervised case, they show that the diffusion maps estimator converges to the regression function as the sample size $n \to \infty$, but at a suboptimal rate.
-As far as we know, there has been no analysis of the test statistic $\wh{T}$ in the random design framework which we study.
-
-There has been much more analysis of the convergence properties of eigenvectors of random graph Laplacians to their continuum limits.
-\citet{belkin07,vonluxburg2008,singer2017, garciatrillos18} show that eigenvalue-eigenvector pairs $(\lambda,v)$ of a graph Laplacian converge to eigenvalue-eigenfunction pairs $(\lambda,\psi)$ of a limiting differential or integral operator.
-\citet{burago2014, shi2015, trillos2019, calder2019, cheng2021} build on these works, by giving finite sample bounds, rates of convergence, and making statements uniform.
-These results justify our intuition that Laplacian eigenmaps, which projects the responses ${\bf Y}$ onto eigenvectors of a graph Laplacian, is approximating a classical spectral projection method, which projects ${\bf Y}$ onto eigenfunctions of the limiting differential operator $\Delta_P$.
-In fact, more formally these results imply that for a fixed number of eigenvectors $K > 0$, the Laplacian eigenmaps estimator converges to its classical counterpart as $n \to \infty$.
-By taking $K \to \infty$ sufficiently slowly with $n$, we can even conclude that Laplacian eigenmaps will be a consistent estimator of $f_0$.
-Unfortunately, the resulting rates of convergence implied by this approach are severely suboptimal, compared to the minimax rates.
-Our analysis showing that Laplacian eigenmaps achieves minimax optimal rates of convergence will use some of the results mentioned above, but will overall proceed by a very different route than the one just considered.
-
-Finally, we point out that there are other ways to use neighborhood graphs, and specifically graph Laplacians, for nonparametric regression.
-For instance, \cite{trillos2020} \textcolor{blue}{(Green 2021)} use the graph Laplacian to induce a penalty over functions $f: \{\bf X\} \to \Reals$.
-The \emph{Laplacian smoothing} estimator $\wh{f}_{\mathrm{LS}}$ is obtained by minimizing the sum of this penalty with a data-fidelity term,
-\begin{equation*}
-\wh{f}_{\mathrm{LS}} := \|Y - f\|_n^2 + \lambda \dotp{L_{n,\varepsilon}f}{f}_n.
-\end{equation*}
-\textcolor{blue}{(Green 2021)} show that the resulting estimator is minimax optimal, but only for $s = 1$ and $d \leq 4$.
-In contrast, we show in this work that the Laplacian eigenmaps estimator is optimal for all $s$ and $d$.
-
-\paragraph{Organization.}
-
-We now outline the structure of the rest of this paper.
-In Section~\ref{sec:setup_main_results}, we formally define the regression problem and estimator we consider.
-We also give some background about minimax regression over Sobolev spaces, and recall the classical spectral projection estimators which achieve minimax rates of convergence.
-In Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity}, we establish our main results regarding the optimality of Laplacian eigenmaps.
-Since thhe Laplacian eigenmaps estimator $\wh{f}$ is defined only at the design points $X_1,\ldots,X_n$, in Section~\ref{sec:out_of_sample} we propose a method for out-of-sample extension of $\wh{f}$, and show that it has optimal out-of-sample error.
-In Section~\ref{sec:experiments} we examine the empirical behavior of Laplacian eigenmaps, and compare it to some natural competitors for nonparametric regression.
-We conclude with some discussion in Section~\ref{sec:discussion}.
-
-\paragraph{Notation.}
-We frequently refer to various classical function classes.
-For a domain $\mc{X}$ with volume form $d\mu$, we let $L^2(\mc{X})$ denote the set of functions $f$ for which $\|f\|_{L^2(\mc{X})}^2 := \int f^2 \,d\mu  < \infty$, and equip $L^2(\mc{X})$ with the norm $\|\cdot\|_{L^2(\mc{X})}$.
-We define $\dotp{f}{g}_P := \int fg\,dP$, and let $L^2(P)$ contain those functions $f$ for which $\|f\|_P^2 := \dotp{f}{f}_P$ is finite.
-Finally, we let $L^2(P_n)$ consist of those ``functions'' $f: \set{X_1,\ldots,X_n} \to \Reals$ for which the empirical norm $\|f\|_{n}^2 := \frac{1}{n}\sum_{i = 1}^{n} \bigl(f(X_i)\bigr)^2 < \infty$.
-When there is no chance of confusion, we will sometimes associate functions in $L^2(P_n)$ with vectors in $\Reals^n$, and vice versa.
-We use $C^k(\mc{X})$ to refer to functions which are $k$ times continuously differentiable in $\mc{X}$, either for some integer $k \geq 1$ or for $k = \infty$.
-We let $C_c^{\infty}(\mc{X})$ represent those functions in $C^{\infty}(\mc{X})$ which are compactly contained in $\mc{X}$.
-We write $\partial f/\partial r_i$ for the partial derivative of $f$ in the $i$th standard coordinate of $\Rd$, and use the multi-index notation $D^{\alpha}f := \partial^{|\alpha|}f/\partial^{\alpha_1}x_1\ldots\partial^{\alpha_d}x_d$ for multi-indices $\alpha \in \Reals^m$.
-
-We write $\|\cdot\|$ for Euclidean, and $d_{\mc{X}}(x',x)$ for the geodesic distance between points $x$ and $x'$ on a manifold $\mc{X}$.
-Then for a given $\delta > 0$, $B(x,\delta)$ is the radius-$\delta$ ball with respect to Euclidean distance, whereas $B_{\mc{X}}(x,\delta)$ is the radius-$\delta$ ball with respect to geodesic distance.
-Letting $T_x(\mc{X})$ be the tangent space at a point $x \in \mc{X}$, we write $B_m(v,\delta) \subset T_x(\mc{X})$ for the radius-$\delta$ ball centered at $v \in T_x(\mc{X})$.
-
-For sequences $(a_n)$ and $(b_n)$, we use the asymptotic notation $a_n \lesssim b_n$ to mean that there exists a number $C$ such that $a_n \leq C b_n$ for all $n$.
-We write $a_n \asymp b_n$ when $a_n \lesssim b_n$ and $b_n \lesssim a_n$.
-On the other hand we write $a_n = o(b_n)$ when $\lim a_n/b_n = 0$, and likewise $a_n = \omega(b_n)$ when $\lim a_n/b_n = \infty$.
-Finally $a \vee b := \max\{a,b\}$ and $a \wedge b := \min\{a,b\}$.
-
-\section{Setup and Background}
-\label{sec:setup_main_results}
-
-In this section, we begin by giving a precise definition of our framework, and the Laplacian eigenmaps methods we study.
-We then review minimax rates for nonparametric regression over Sobolev spaces.
-We pay special attention to the classical spectral projection methods which achieve these rates, since such methods are closely connected to Laplacian eigenmaps.
-
-\subsection{Nonparametric regression with random design}
-\label{sec:regression_laplacian_eigenmaps}
-
-We will operate in the usual setting of nonparametric regression with random design, in which we observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$.
-The design points $X_1,\ldots,X_n$ are sampled from a distribution $P$ with support $\mc{X} \subseteq \Rd$, and the responses follow the signal plus noise model
-\begin{equation}
-\label{eqn:model}
-Y_i = f_0(X_i) + w_i,
-\end{equation}
-with regression function $f_0: \mc{X} \to \Reals$, and $w_i \sim N(0,1)$ independent Gaussian noise.
-For simplicity we will assume throughout that the noise has unit-variance, but all of our results extend in a straightforward manner to the case where the variance is equal to a known positive value.
-
-We now formulate two models, which differ in the assumed nature of the support $\mc{X}$ of the design distribution $P$: the \emph{flat Euclidean} and \emph{manifold} models.
-
-\begin{definition}[Flat Euclidean model]
-	\label{def:model_flat_euclidean}
-	The data $(X_1,Y_1),\ldots,(X_n,Y_n)$ are sampled according to~\eqref{eqn:model}.
-The support $\mc{X}$ of the design distribution $P$ is an open, connected, and bounded subset of $\Rd$, with Lipschitz boundary.
-The distribution $P$ admits a Lipschitz density $p$ with respect to the $d$-dimensional Lebesgue measure $\nu$, which is bounded away from $0$ and $\infty$,
-	\begin{equation*}
-	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty, \quad \textrm{for all $x \in \mc{X}$.}
-	\end{equation*}
-\end{definition}
-
-In the following, we recall that the injectivity radius of a $m$-dimensional Riemannian manifold $\mc{X}$ is the maximum value of $\delta$ such that the exponential map $\exp_x: B_m(0,\delta) \subset T_x(\mc{X}) \to B_{\mc{X}}(x,\delta) \subset \mc{X}$ is a diffeomorphism for all $x \in \mc{X}$.
-\begin{definition}[Manifold model]
-	\label{def:model_manifold}
-	The data $(X_1,Y_1),\ldots,(X_n,Y_n)$ are sampled according to~\eqref{eqn:model}. 
-	The support $\mc{X}$ of the design distribution $P$ is a closed, connected, smooth and boundaryless Riemannanian manifold embedded in $\Rd$, of intrinsic dimension $1 \leq m < d$.
-The injectivity radius of $\mc{X}$ is lower bounded by a positive constant $i_0 > 0$.
-The design distribution $P$ admits a Lipschitz density $p$ with respect to the volume form $d\mu$ induced by the Riemannian structure of $\mc{X}$, which is bounded away from $0$ and $\infty$,
-	\begin{equation*}
-	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty, \quad \textrm{for all $x \in \mc{X}$.}
-	\end{equation*}
-\end{definition}
-
-Finally, at various points we will have to assume that the density $p$ also displays different types of higher-order regularity, beyond Lipschitz continuity.
-When such assumptions are necessary, we always will state them explicitly.
-
-\subsection{Laplacian eigenmaps}
-We now formally define the estimator and test statistic we study.
-Both are derived from eigenvectors of a graph Laplacian.
-For a positive, symmetric kernel $\eta: [0,\infty) \to [0,\infty)$, and a radius parameter $\varepsilon > 0$, let $G = ([n],W)$ be the neighborhood graph formed over the design points $\{X_1,\ldots,X_n\}$, with a weighted edge $W_{ij} = \eta(\|X_i - X_j\|/\varepsilon)$ between vertices $i$ and $j$.
-Then the
-\emph{neighborhood graph Laplacian} $L_{n,\varepsilon}: \Reals^n \to \Reals$ is defined by its action on vectors $u \in \Reals^n$ as
-\begin{equation}
-\label{eqn:neighborhood_graph_laplacian}
-\bigl(L_{n,\varepsilon}u\bigr)_i := \frac{1}{n\varepsilon^{2 + \mathrm{dim}(\mc{X})}} \sum_{j = 1}^{n} \bigl(u_i - u_j\bigr) \eta\biggl(\frac{\|X_i - X_j\|}{\varepsilon}\biggr).
-\end{equation}
-(Here $\mathrm{dim}(\mc{X})$ stands for the dimension of $\mc{X}$.
-It is equal to $d$ under the assumptions of Model~\ref{def:model_flat_euclidean}, and equal to $m$ under the assumptions of Model~\ref{def:model_manifold}.
-The pre-factor $(n\varepsilon^{2 + \mathrm{dim}(\mc{X})})^{-1}$ is purely for convenience in taking limits as $n \to \infty, \varepsilon \to 0$).
-Written in standard coordinates we have $(n\varepsilon^{\dim(\mc{X}) + 2}) \cdot L_{n,\varepsilon} = D - W$, where $D \in \Reals^{n \times n}$ is the diagonal degree matrix, $D_{ii} = \sum_{i = 1}^{n} W_{ij}$.
-
-The graph Laplacian is a positive semi-definite matrix, and admits the eigendecomposition $L_{n,\varepsilon} = \sum_{k = 1}^{n} \lambda_k v_k v_k^{\top}$, where for each $k = 1,\ldots,n$ the eigenvalue-eigenvector pair $(\lambda_k,v_k)$ satisfies
-\begin{equation*}
-\frac{1}{n}L_{n,\varepsilon}v_k = \lambda_k v_k, \quad \|v_k\|_n^2 = 1.
-\end{equation*}
-We will assume without loss of generality that each eigenvalue $\lambda$ of $L_{n,\varepsilon}$ has algebraic multiplicity $1$, and so we can index the eigenpairs $(\lambda_1,v_1),\ldots,(\lambda_n,v_n)$ in ascending order of eigenvalue, $0 = \lambda_1 < \ldots < \lambda_n$.
-
-The Laplacian eigenmaps estimator $\wh{f}$ simply projects the response vector ${\bf Y}$ onto the first $K$ eigenvectors of $L_{n,\varepsilon}$: letting $V_K \in \Reals^{n \times K}$ be the matrix with columns $v_1,\ldots,v_K$, we have that
-\begin{equation}
-\label{eqn:laplacian_eigenmaps_estimator}
-\wh{f} := \sum_{k = 1}^{K} \dotp{Y}{v_k}_{n} v_k = \frac{1}{n} V_K V_K^{\top} Y.
-\end{equation} 
-Thus $\wh{f}$ is equivalently a vector in $\Reals^n$, or a function in $L^2(P_n)$. If $\wh{f}$ is a reasonable estimate of $f_0$, then the Laplacian eigenmaps test statistic
-\begin{equation}
-\label{eqn:laplacian_eigenmaps_test}
-\wh{T} := \|\wh{f}\|_n^2
-\end{equation}
-is in turn a reasonable estimate of $\|f_0\|_{P}^2$, and can be used to distinguish whether or not $f_0 = 0$.
-
-It may be helpful to comment briefly on the term ``Laplacian eigenmaps'', which we use a bit differently than is typical in the literature.
-Laplacian eigenmaps typically refers to an algorithm for embedding, which maps each design point $X_1,\ldots,X_n$ to $\Reals^K$ according to $X_i \mapsto (v_{1,i}, \ldots, v_{K,i})$.
-Viewing this embedding as a feature map, we can then interpret the estimator $\wh{f}$ as the least-squares solution to a linear regression problem with responses $Y_1,\ldots,Y_n$ and features $v_1,\ldots,v_K$.
-Often, the Laplacian eigenmaps embedding is viewed as a tool for dimensionality reduction, wherein it is implicitly assumed that $K$ is much smaller than $d$.
-We will neither explicitly nor implicitly take $K < d$; indeed, the embedding perspective is not particularly illuminating in what follows, and we do not henceforth make reference to it.
-Instead, we use ``Laplacian eigenmaps'' to directly refer to the estimator $\wh{f}$ or test statistic $\wh{T}$.
-
-\subsection{Sobolev Classes}
-\label{sec:sobolev}
-We now review the definition of Sobolev classes, dividing our discussion into two cases---the flat Euclidean case and the manifold case, corresponding respectively to Models~\ref{def:model_flat_euclidean} and Model~\ref{def:model_manifold}.
-
-\paragraph{Flat Euclidean case.}
-Take $\mc{X}$ to be a open, connected and bounded set with Lipschitz boundary, as in Model~\ref{def:model_flat_euclidean}.
-Recall that for a given multiindex $\alpha \in \mathbb{N}^d$, a function $f$ is \emph{$\alpha$-weakly differentiable} if there exists some $h \in L^1(\mc{X})$ such that
-\begin{equation*}
-\int_{\mc{X}} h g = (-1)^{|\alpha|} \int_{\mc{X}} f D^{\alpha}g, \quad \textrm{for every $g \in C_c^{\infty}(\mc{X})$.}
-\end{equation*}
-If such a function $h$ exists, it is the $\alpha$th weak partial derivative of $f$, and denoted by $D^{\alpha}f := h$.
-
-\begin{definition}[Sobolev space on an open set]
-	\label{def:sobolev_space}
-	For an integer $s \geq 1$, a function $f \in L^2(\mc{X})$ belongs to the Sobolev space $H^s(\mc{X})$ if for all $|\alpha| \leq s$, the weak derivatives $D^{\alpha}f$ exists and satisfy $D^{\alpha}f \in L^2(\mc{X})$.
-The $j$th order semi-norm for $f \in H^s(\mc{X})$ is $|f|_{H^j(\mc{X})} := \sum_{|j| = s}\|D^{\alpha}f\|_{\Leb^2(\mc{X})}$, and the corresponding norm
-	\begin{equation*}
-	\|f\|_{H^s(\mc{X})}^2 := \|f\|_{\Leb^2(\mc{X})}^2 + \sum_{j = 1}^{s} |f|_{H^j(\mc{X})}^2,
-	\end{equation*}
-	induces a ball
-	\begin{equation*}
-	H^s(\mc{X};M) := \bigl\{f \in H^s(\mc{X}): \|f\|_{H^s(\mc{X})} \leq M\bigr\}.
-	\end{equation*} 
-\end{definition}
-We note that $H^s(\mc{X})$ is the completion of $C^{\infty}(\mc{X})$ with respect to the $\|\cdot\|_{H^s(\mc{X})}$ norm, so that $C^{\infty}(\mc{X})$ is dense in $H^s(\mc{X})$.
-
-\paragraph{Manifold case.}
-
-There are several equivalent ways to define Sobolev spaces on a compact, smooth, $m$-dimensional Riemannian manifold embedded in $\Reals^d$.
-We will stick with a definition that parallels our setup in the flat Euclidean setting as much as possible.
-To do so, we first recall the notion of partial derivatives on a manifold, which are defined with respect to a local coordinate system.
-Letting $r_1,\ldots,r_m$ be the standard basis of $\Reals^m$, for a given chart $(\phi,U)$ (meaning an open set $U \subseteq \mc{X}$, and a smooth mapping $\phi: U \to \Reals^m$) we write $\phi =: (x_1,\ldots,x_m)$ in local coordinates, meaning $x_i = r_i \circ \phi$.
-Then the partial derivative $\partial f/\partial x_i$ of a function $f$ with respect to $x_i$ at $x \in U$ is
-\begin{equation*}
-\frac{\partial f}{\partial x_i}(x) := \frac{\partial(f \circ \phi^{-1})}{\partial r_i}\bigl(\phi(x)\bigr).
-\end{equation*}
-The right hand side should be interpreted in the weak sense of derivative.
-As before, we use the multi-index notation $D^{\alpha}f := \partial^{|\alpha|}f/\partial^{\alpha_1}x_1\ldots\partial^{\alpha_m}x_m$.
-
-\begin{definition}[Sobolev space on a manifold]
-	\label{def:sobolev_space_manifold}
-	A function $f \in \Leb^2(\mc{X})$ belongs to the Sobolev space $H^{s}(\mc{X})$ if for all $\abs{\alpha} \leq s$, the weak derivatives $D^{\alpha}f$ exist and satisfy  $D^{\alpha}f \in \Leb^2(\mc{X})$.
-The $j$th order semi-norm $|f|_{H^j(\mc{X})}$, the norm $\|f\|_{H^s(\mc{X})}$, and the ball $H^s(\mc{X};M)$ are all defined as in Defintion~\ref{def:sobolev_space}.
-\end{definition}
-The partial derivatives $D^{\alpha}f$ will depend on the choice of local coordinates, and so will the resulting Sobolev norm~$\|f\|_{H^s(\mc{X})}$.
-However, regardless of the choice of local coordinates the resulting norms will be equivalent\footnote{Recall that norms $\|\cdot\|_1$ and $\|\cdot\|_2$ on a space $\mc{F}$ are said to be equivalent if there exist constants $c$ and $C$ such that
-\begin{equation*}
-c \|f\|_1 \leq \|f\|_2 \leq C \|f\|_2 \quad \textrm{for all} f \in \mc{F}.
-\end{equation*}} 
-and so the ultimate Sobolev space $H^s(\mc{X})$ is independent of the choice of local coordinates.
-
-% AG: This was my first attempt at defining Sobolev spaces on manifolds, which I keep as a comment for now, and plan to delete.
-%
-% We stick to an approach that builds on the definition of Sobolev spaces in the flat Euclidean case.
-Specifically, let $\{\phi_{\beta}, U_{\beta}\}$ be a $C^{\infty}$ atlas of $\mc{X}$, and let $\{h_\beta\}$ be a partition of unity subordinate to $\{(\phi_{\beta}), U_{\beta}\}$.\footnote{Meaning $\mathrm{supp}(h_{\beta}) \subseteq U_{\beta}$ for each $\beta$.} For a function $f \in C^{\infty}(\mc{X})$, define the order-$s$ Sobolev norm to be
-%\begin{equation*}
-%\|f\|_{H^s(\mc{X})}^2 := \sum_{\beta} \|(fh_{\beta}) \circ %\phi_\beta^{-1}\|_{H^s(\phi(U_\beta))}^2.
-%\end{equation*}
-%Then the order-$s$ Sobolev space $H^s(\mc{X})$ is defined to be the completion of $C^{\infty}(\mc{X})$ with respect to the norm $\|\cdot\|_{H^s(\mc{X})}$.
-When the manifold $\mc{X}$ is smooth, as we assume in Model~\ref{def:model_manifold}, then this construction is independent of the choice of atlas $\{\phi_{\beta}, U_{\beta}\}$.
-
-\paragraph{Boundary conditions.}
-In the flat Euclidean model (Model~\ref{def:model_flat_euclidean}), in order to show that Laplacian eigenmaps is optimal over $H^{s}(\mc{X})$ for $s > 1$ we will need to assume that the regression function $f_0$ satisfies some boundary conditions.
-In particular, we will assume that $f_0$ is zero-trace.
-\begin{definition}[Zero-trace Sobolev space]
-	\label{def:zero_trace_sobolev_space}
-	The \emph{order-s zero-trace Sobolev space} $H_0^s(\mc{X})$ is the closure of $C_c^\infty(\mc{X})$ with respect to $\|\cdot\|_{H^s(\mc{X})}$ norm.
-That is, $f \in H_0^s(\mc{X})$ if $f \in H^s(\mc{X})$ and additionally there exists a sequence $f_1,f_2,\ldots$ of functions in $C_c^{\infty}(\mc{X})$ such that
-	\begin{equation*}
-	\lim_{k \to \infty}\|f_k - f\|_{H^s(\mc{X})} = 0.
-	\end{equation*}
-	The normed ball $H_0^{s}(\mc{X};M) := H_0^{s}(\mc{X}) \cap H^{s}(\mc{X};M)$.
-\end{definition}
-The zero-trace condition can be made more concrete when $f \in C^\infty(\mc{X})$, since we can then speak of the pointwise behavior of $f$ and its derivatives.
-Letting $\partial/(\partial {\bf n})$ be the partial derivative operator in the direction of the vector $\mathbf{n}$ normal to the boundary of $\mc{X}$, then the zero-trace condition implies that $\partial^{k}f/\partial{\bf n}^k(x) = 0$ for each $k = 0,\ldots,s - 1$, and for all $x \in \partial\mc{X}$.
-
-We now explain why Laplacian eigenmaps should be optimal only when $f_0$ satisfies certain boundary conditions.
-Let $(\lambda_1(\Delta_P),\psi_1),(\lambda_2(\Delta_P),\psi_2),\ldots$ be the solutions to the weighted Laplacian eigenvector equation with Neumann boundary conditions,
-\begin{equation}
-\label{eqn:laplacian_eigenfunction_neumann}
-\Delta_P\psi_k = \lambda_{k}(\Delta_P) \psi_k, \quad \frac{\partial}{\partial {\bf n}}\psi_k = 0~~\textrm{on $\partial \mc{X}$}.
-\end{equation}
-As we have already alluded to, it is known~\citep{garciatrillos18} that each graph Laplacian eigenpair ($\lambda_k$,$v_k$) converges to a corresponding solution ($\lambda_k(\Delta_P)$, $\psi_k$) of~\eqref{eqn:laplacian_eigenfunction_neumann}.
-Thus it is relevant to consider which Sobolev functions $f \in H^s(M)$ we can reconstruct using the eigenfunctions $\psi_1,\psi_2,\ldots$.
-To that end, we introduce the spectrally defined Sobolev space
-\begin{equation}
-\label{eqn:spectral_sobolev_space}
-\mc{H}^s(\mc{X}) := \biggl\{f \in L^2(\mc{X}): \sum_{k = 1}^{\infty} \bigl[\dotp{f}{\psi_k}_P\bigr]^2 \cdot\bigl[\lambda_k(\Delta_P)\bigr]^s < \infty \biggr\}.
-\end{equation}
-Under the conditions $p \in C^{\infty}(\mc{X})$ and $\partial\mc{X} \in C^{1,1}$, \cite{dunlop2020} show the strict inclusion $\mc{H}^{2s}(\mc{X}) \subset H^{2s}(\mc{X})$.
-More precisely, they show that for any $s > 0$,
-\begin{equation}
-\label{eqn:spectral_sobolev_space_equivalence}
-\mc{H}^{2s}(\mc{X}) = \biggl\{f \in H^{2s}(\mc{X}): \frac{\partial \Delta_P^rf}{\partial {\bf n}} = 0~\textrm{on}~\partial\mc{X},~~\textrm{for all $0 \leq r \leq s - 1$} \biggr\},
-\end{equation}
-and for any $s \geq 0$, $\mc{H}^{2s + 1}(\mc{X}) = \mc{H}^{2s}(\mc{X}) \cap H^{2s + 1}(\mc{X})$.
-If $P$ is uniform on $\mc{X}$, then~\eqref{eqn:spectral_sobolev_space} means that a Sobolev function $f \in H^{s}(\mc{X})$ additionally belongs to $\mc{H}^{s}(\mc{X})$ only if all its odd lower order derivatives vanish at the boundary $\partial\mc{X}$.
-
-The bottom line is that the eigenvectors $v_k$ of the graph Laplacian accurately approximate only those functions $f \in H^{s}(\mc{X})$ which satisfy some additional boundary conditions.
-Although the zero-trace boundary condition is more restrictive than~\eqref{eqn:spectral_sobolev_space_equivalence}---since it also requires that derivatives of even-order be equal to $0$---the point is that some kind of boundary condition will be necessary in order to obtain optimal rates.
-We will stick to the zero-trace condition, since it greatly eases some of the steps in our proofs.
-
-On the other hand, in the manifold model (Model~\ref{def:model_manifold}) the domain $\mc{X}$ is without boundary---precisely, every point $x \in \mc{X}$ has a neighborhood that is homeomorphic to an open set in $\Reals^m$, for instance the open ball $B(x,\delta)$ for any $\delta$ smaller than the injectivity radius $i_0$---and so boundary conditions are irrelevant.
-
-\subsection{Minimax Rates and Spectral Series Methods}
-\label{subsec:minimax_rates_sobolev}
-We now review the minimax estimation and goodness-of-fit testing rates over Sobolev balls.
-We will pay special attention to certain classical spectral projection methods which achieve these rates.
-This is because, as we have already discussed, classical spectral projection methods are very related to Laplacian eigenmaps.
-
-\paragraph{Estimation rates over Sobolev balls.}
-In the estimation problem, we ask for an estimator---formally, a Borel measurable function $\wh{f}$ that maps from $\mc{X}$ to $\Reals$---which is close to the regression function $f_0$ with respect to the squared norm $\|\wh{f} - f_0\|_{P}^2$.
-Under Model~\ref{def:model_flat_euclidean}, the minimax estimation rate over the order-$s$ Sobolev ball is
-\begin{equation}
-\label{eqn:sobolev_space_minimax_estimation_rate}
-\inf_{\wh{f}} \sup_{f_0} \Ebb\|\wh{f} - f_0\|_P^2 \asymp M^2(M^2n)^{-2s/(2s + d)};
-\end{equation}
-here infimum taken over all estimators $\wh{f}$, and the supremum over all $f_0 \in H^1(\mc{X};M)$ (first-order case) or $f_0 \in H_0^s(\mc{X};M)$ (higher-order case), and we assume $n^{-1/2} \lesssim M$.
-
-The lower bound in~\eqref{eqn:sobolev_space_minimax_estimation_rate} is due to~\citep{stone1980} (at least for the case of $M$ constant, as is most typically considered).
-The upper bound can be certified by a particular spectral projection estimator $\wt{f}$,
-\begin{equation}
-\label{eqn:spectral_series_estimator}
-\wt{f} := \sum_{k = 1}^{K} \dotp{Y}{\psi_k}_n \psi_k,
-\end{equation}
-where $\psi_k$ are the eigenfunctions of $\Delta_P$ defined in~\eqref{eqn:laplacian_eigenfunction_neumann}.
-The optimality of spectral projection estimators over Sobolev type spaces is generally well-understood.
-For instance, see \cite{tsybakov08,johnstone2011,gine16}, who work in the Gaussian sequence model and show that analogous estimators are optimal over Sobolev ellipsoids.
-However, we have not found an analysis of the specific estimator~\eqref{eqn:spectral_series_estimator} under Model~\ref{def:model_flat_euclidean}, and so for completeness we state the result in the following proposition.
-\begin{proposition}
-	\label{prop:spectral_series_estimation}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $\partial \mc{X} \in C^{1,1}$, that $p \in C^{\infty}(\mc{X})$, and that $f_0 \in H^1(\mc{X};M)$ (first-order case) or $f_0 \in H_0^s(\mc{X};M)$ for some $s > 1$ (higher-order case).
-Then there exists a constant $C$ which does not depend on $f_0,M$ or $n$ such that the following statement holds: if the spectral projection estimator $\wt{f}$ is computed with parameter $K = \floor{M^2n}^{d/(2s + d)}$, then
-	\begin{equation*}
-	\Ebb\bigl[\|\wt{f} - f_0\|_P^2\bigr] \leq C \min\bigl\{M^2(M^2n)^{-2s/(2s + d)}, M^2\bigr\}
-	\end{equation*}
-\end{proposition}
-The proof of Proposition~\ref{prop:spectral_series_estimation}, along with proofs of all of our results, can be found in the appendix.
-It is worth mentioning some aspects of the analysis here, because it sets the stage for the strategy we will use to analyze Laplacian eigenmaps.
-
-In particular, three essential facts are needed to establish Proposition~\ref{prop:spectral_series_estimation}. 
-\begin{enumerate}
-	\item The continuous embedding of $H_0^s(\mc{X})$ into $\mc{H}^s(\mc{X})$---recall the latter is defined in~\eqref{eqn:spectral_sobolev_space}---which is a consequence of the zero-trace condition, the conditions on $\partial \mc{X}$ and $p$, and~\eqref{eqn:spectral_sobolev_space_equivalence}.
-	\item Weyl's Law, which gives the asymptotic scaling of eigenvalues $\lambda_{k}(\Delta_P) \asymp k^{2/d}$, and allows us to properly control the bias induced by spectral projection.
-	\item A local version of Weyl's law, which gives an estimate on $\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2$, and allows us to appropriately control the difference $\dotp{f_0}{\psi_k}_n - \dotp{f_0}{\psi_k}_P$  between the empirical and population Fourier coefficients.
-\end{enumerate}
-With these facts in hand the proof of Proposition~\ref{prop:spectral_series_estimation} follows from calculations standard to analysis of the Gaussian sequence model.
-Our analysis of Laplacian eigenmaps will depend on analogues to the first and second of these facts, with the space $\mc{H}^s(\mc{X})$ and eigenvalues $\lambda_k(\Delta_P)$ replaced by alternatives suitably defined with respect to the neighborhood graph Laplacian $L_{n,\varepsilon}$.
-
-\paragraph{Goodness-of-fit testing rates over Sobolev balls.}
-In the goodness-of-fit testing problem, we ask for a test function---formally, a Borel measurable function $\phi$ that takes values in $\{0,1\}$--- which can distinguish between the hypotheses
-\begin{equation}
-\mathbf{H}_0: f_0 = f_0^{\ast}, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{f_0^{\ast}\}.
-\end{equation} 
-Typically, the null hypothesis $f_0 = f_0^{\ast} \in \mc{F}$ reflects the absence of interesting structure, and $\mc{F} \setminus  \{f_0^{\ast}\}$ is a set of smooth departures from this null.
-To fix ideas, as in \citet{ingster2009} we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\ast} = 0$ and $\mc{F} = \mc{H}^s(\mc{X};M)$ is a Sobolev ball.
-This is without loss of generality since our test statistic and its analysis are easily modified to handle the case when $f_0^{\ast}$ is not $0$, by simply subtracting $f_0^{\ast}(X_i)$ from each observation $Y_i$.
-
-The Type I error of a test $\phi$ is $\mathbb{E}_0[\phi]$, and if $\mathbb{E}_0[\phi] \leq a$ for a given $a \in (0,1)$ we refer to $\phi$ as a level-$a$ test\footnote{We reserve the more common symbol $\alpha$ for multi-indices, so as to avoid confusion}.
-The worst-case risk of $\phi$ over $\mc{F}$ is
-\begin{equation*}
-R_n(\phi, \mc{F}, \epsilon) := \sup\Bigl\{\mathbb{E}_{f_0}[1 - \phi]: f_0 \in \mc{F}, \|f_0\|_{P} > \epsilon\Bigr\},
-\end{equation*}
-and for a given constant $b \in (0,1)$, the minimax critical radius $\epsilon_n(\mc{F})$ is the smallest value of $\epsilon$ such that some level-$a$ test has worst-case risk of at most $b$. Formally,
-\begin{equation*}
-\epsilon_n(\mc{F}) := \inf\Bigl\{\epsilon > 0: \inf_{\phi} R_n(\phi,\mc{F},\epsilon) \leq b \Bigr\},
-\end{equation*} 
-where in the above the infimum is over all level-$a$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Clearly, the minimax critical radius $\epsilon_n(\mc{F})$ depends on $a$ and $b$.
-However, we adopt the typical convention of treating $\alpha,b \in (0,1)$ and  as small but fixed positive constants; hence they will not affect the testing error rates, and we suppress them notationally.} We will refer to the rate at which the squared critical radius $\epsilon_n(\mc{F})^2$ tends to $0$ as the minimax testing rate.
-
-Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing rate is much smaller than the minimax estimation rate.
-\citet{ingster2009} give the minimax testing rate in the special case where $\mc{X} = [0,1]^d$ and $P$ is uniform, and we restate their result in the terms and notation of our paper.
-\begin{theorem}[Theorem 1 of \citet{ingster2009}]
-	\label{thm:spectral_series_testing}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $\mc{X} = [0,1]^d$, $P$ is the uniform distribution on $[0,1]^d$, and $f_0 \in H_0^s(\mc{X})$. Then 
-	\begin{equation}
-	\label{eqn:sobolev_space_testing_critical_radius}
-	\epsilon_n^2\bigl(H_0^s(P;1)\bigr) \asymp n^{-4s/(4s + d)}~~\textrm{for $1 \leq d < 4s$.}
-	\end{equation}
-\end{theorem}
-The analysis used by \citet{ingster2009} to show the upper bound in~\eqref{eqn:sobolev_space_testing_critical_radius} relies on a similar trio of facts as used in the proof of Proposition~\ref{prop:spectral_series_estimation}.
-It is otherwise reminiscent of calculations made in the Gaussian sequence model, which can be found in~\cite{ingster2012}.
-This analysis can be straightforwardly adapted to handle design distributions $P$ that satisfy the conditions of Model~\ref{def:model_flat_euclidean}, or to handle the case where $M$ is not $1$.
-Finally, the test \citet{ingster2009} use to certify the upper bound in~\eqref{eqn:sobolev_space_testing_critical_radius} is implicitly a spectral projection method: the test statistic is the $L^2(P_n)$ norm of a particular spectral projection estimator.\footnote{\cite{ingster2009} project the responses onto a the span of trigonometric basis functions.
-Since $P$ is uniform on the unit cube, such functions are eigenfunctions of $\Delta_P$ when the right boundary conditions are imposed.}
-
-A major difference between testing and estimation over Sobolev spaces is the requirement that $4s > d$.
-When $4s \leq d$, the functions in $H^s(\Xset)$ are very irregular.
-Crucially, $H^s(\Xset)$ no longer continuously embeds into $\Leb^4(\Xset)$ when $4s \leq d$, and test statistics using the $L^2(P_n)$ norm are no longer guaranteed to have finite variance.\footnote{Note that this will not affect the analysis for estimation, because for estimation we only need to control the first two moments of $f_0$.} In fact, we have not seen any analysis which describes the minimax rate for nonparametric regression testing over Sobolev spaces in the $4s \leq d$ regime.
-However, if one explicitly assumes that $f_0 \in \Leb^4(\mc{X})$, then the critical radius is characterized by the dimension-free rate $\epsilon_n^2(\Leb^4(\mc{X};M)) \asymp Mn^{-1/2}$.\footnote{As a sanity check note that this is strictly worse than the rate in~\eqref{eqn:sobolev_space_testing_critical_radius}.
-In other words, whenever $4s > d$, the embedding $H^s(\mc{X}) \subseteq L^4(\mc{X})$ never yields a tight upper bound.} As we discuss after our first main theorem regarding testing with Laplacian eigenmaps (Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}), this rate is achievable by a test based on $\wh{T}$.
-
-\paragraph{Manifold setup.}
-Under Model~\ref{def:model_manifold}, both~\eqref{eqn:sobolev_space_minimax_estimation_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced everywhere by the intrinsic dimension $m$.
-The estimator $\wt{f}$ and a test using the statistic $\wt{T}$---with $\psi_1,\psi_2,\ldots$ now the eigenfunctions of the manifold weighted Laplace-Beltrami operator $\Delta_P$---achieve the optimal estimation and testing rates.
-This is because each of the three facts mentioned after Proposition~\ref{prop:spectral_series_estimation} have analogues when the domain is a smooth manifold (See~\citet{hendriks1990}, who analyzes a spectral projection density estimator, for details.)
-
-\paragraph{In-sample mean squared error.}
-As mentioned in our introduction, roughly speaking one of our main conclusions is that the Laplacian eigenmaps estimator $\wh{f}$ is minimax rate-optimal.
-It is worth being clear about what we do and do not mean by this statement.
-We do not mean that the estimator $\wh{f}$ will match the upper bound given in~\eqref{eqn:sobolev_space_minimax_estimation_rate}, since such a statement does not make sense when the estimator is defined only at the random design points $X_1,\ldots,X_n$.
-Instead we will measure loss using the squared $L^2(P_n)$ error.
-In Section~\ref{sec:out_of_sample} we show that an extension of $\wh{f}$ defined over all $\mc{X}$ has $L^2(P)$ error comparable to the $L^2(P_n)$ error of $\wh{f}$.
-We also believe that in the random design setting we work in, simple arguments will imply that $L^2(P_n)$ risk has the same minimax rate of convergence as $L^2(P)$ risk.
-We sketch such an argument in Section~\ref{sec:out_of_sample}, but do not further pursue the details.
-
-Additionally, we will not actually measure accuracy using the expectation of the loss.
-Rather, we will give a high-probability bound on $\|\cdot\|_n^2$.
-For instance, when $f_0 \in H^1(\mc{X};1)$, we will show that with probability $1 - \delta$ the loss $\|\wh{f} - f_0\|_n^2 \leq C_{\delta} n^{-2/(2 + d)}$, for a constant $C_{\delta}$ that depends on $\delta$ but not on $f_0$ or $n$.
-Thus we give an upper bound on the $(1 - \delta)$th quantile of $\|\wh{f} - f_0\|_n^2$, rather than an upper bound on its expectation.
-We explain the reason for this in Section~\ref{sec:minimax_optimal_laplacian_eigenmaps}.
-We also show that if $f_0$ is bounded in a larger norm---for instance, if it is H\"{o}lder rather than Sobolev smooth---then we can obtain bounds on the expected $L^2(P_n)$ loss.
-
-There is one other subtlety introduced by the use of in-sample mean squared error.
-Technically speaking, elements $f \in H^s(\mc{X})$ are equivalence classes, defined only up to a set of measure zero.
-Thus one cannot speak of the pointwise evaluation $f_0(X_i)$, as we do by defining our target of estimation to be $f_0(X_i)$, $i=1,\ldots,n$, until one selects \emph{representatives}.
-When $s > d/2$, every element $f$ of $H^s(\mc{X})$ admits a continuous version $f^{\ast}$, and as is standard we set this to be our favored representative.
-When $s \leq d/2$, some elements in $H^s(\mc{X})$ do not have any continuous version; however they admit a \emph{quasi-continuous} version \citep{evans15} known as the \emph{precise representative}, and we use this representative.
-To be clear, however, it does not really matter which representative we choose.
-Since all versions agree except on a set of measure zero, and since $P$ is absolutely continuous with respect to Lebesgue measure (in Model~\ref{def:model_flat_euclidean}) or the volume form $d\mu$ (in Model~\ref{def:model_manifold}), with probability $1$ any two versions $g_0, h_0 \in f_0$ will satisfy $g_0(X_i) = h_0(X_i)$ for all $i = 1,\ldots,n$.
-The bottom line is that we can use the notation $f_0(X_i)$ without fear of ambiguity or confusion.
-
-Finally, we note that for testing none of these comments are relevant.
-We will show that our test has small worst case risk whenever $\epsilon \gtrsim \epsilon_n(H_0^s(\mc{X};M))$, thus establishing that it is a minimax optimal test in the usual sense.
+\input{sections/introduction}
+\input{sections/preliminaries}
 
@@
 \label{sec:first_order_sobolev_classes}
-We begin by assuming $f_0 \in H^1(\mc{X})$.
+We begin by assuming $f_0 \in H^1(\mc{X}; M)$.
 We show that $\wh{f}$ and a test based on $\wh{T}$ are minimax optimal, for all values of $d$, and under no additional assumptions (beyond those of Model~\ref{def:model_flat_euclidean}) on the data generating process, i.e.
@@
 
-\paragraph{Estimation.} When the kernel $\eta$, graph radius $\varepsilon$, and number of eigenvectors $K$ are chosen appropriately, we show in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that the estimator $\wh{f}$ achieves the minimax rate over $H^1(\mc{X};M)$, when error is measured in squared $L^2(P_n)$ norm.
+\paragraph{Estimation.} When the kernel $\eta$, graph radius $\varepsilon$, and number of eigenvectors $K$ are chosen appropriately, we show in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that the estimator $\wh{f}$ achieves the minimax rate over $H^1(\mc{X};M)$, when the error is measured in squared $L^2(P_n)$ norm.
 \begin{enumerate}[label=(K\arabic*)]
@@
 	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H^1(\mc{X},M)$.
-There are constants $c,C$ and $N$ (not depending on $f_0$, $M$ or $n$), such that the following statement holds for all $n \geq N$ and any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_fo}, then
+There are constants $c,C$ and $N$ (not depending on $f_0$, $M$ or $n$), such that the following statement holds for all $n \geq N$ and any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_fo}, then
 	\begin{equation}
@@
 \end{theorem}
-From~\eqref{eqn:laplacian_eigenmaps_estimation_fo} it follows immediately that with high probability $\|\wh{f} - f_0\|_n^2 \lesssim M^2(M^2n)^{-2/(2 + d)}$ whenever $n^{-1/2} \lesssim M \lesssim n^{1/d}$.
+From~\eqref{eqn:laplacian_eigenmaps_estimation_fo} it follows immediately that with constant probability $\|\wh{f} - f_0\|_n^2 \lesssim M^2(M^2n)^{-2/(2 + d)}$ whenever $n^{-1/2} \lesssim M \lesssim n^{1/d}$.
 
@@
 On the other hand, as we will see in Section~\ref{subsec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K$ is of at least the same order as the continuum eigenvalue $\lambda_K(\Delta_P)$.
-Finally, we choose $K = (M^2n)^{2d/(2 + d)}$ (when possible) to optimally trade-off bias and variance.
-	\item We note that the ranges~\eqref{eqn:radius_fo} and~\eqref{eqn:eigenvector_estimation_fo} depend on quantities, such as the dimension $d$ and radius of the Sobolev ball $M$, which are usually unknown.
+Finally, we choose $K = (M^2n)^{d/(2 + d)}$ (when possible) to optimally trade-off bias and variance.
+	\item We note that the ranges~\eqref{eqn:radius_fo} and~\eqref{eqn:eigenvector_estimation_fo} depends on the radius of the Sobolev ball $M$, which is usually unknown.
 In practice, one typically tunes hyper-parameters by sample-splitting or cross-validation.
@@
 This choice of threshold $t_{a}$ guarantees that $\varphi$ is a level-$a$ test.
-Moreover, when $\varepsilon$ and $K$ are chosen appropriately, the test $\varphi$ has negligible Type II error against alternatives separated from the null by at least $\|f_0\|_{P}^2 \gtrsim M^2(M^2n)^{-4/(4 + d)}$, whenever $d < 4$.
+As we show in Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}, when $d < 4$, $\varepsilon$ and $K$ are chosen appropriately, and the alternative $f_0$ is separated from $0$ by at least (a constant times) the minimax critical radius $M^2(M^2n)^{-4/(4 + d)}$, the test $\varphi$ has Type II error of at most $b$.
 
@@
 	\label{asmp:parameters_testing_fo}
-	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy~\eqref{eqn:radius_fo}. Additionally,
+	The graph radius $\varepsilon$ satisfies~\eqref{eqn:radius_fo}, and the number of eigenvectors 
 	\begin{equation}
@@
 Suppose additionally $f_0 \in H^1(\mc{X},M)$, and that $d < 4$.
-Then there exist constants $C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_fo}, and if $f_0$ satisfies
+Then there exist constants $C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_fo}, and if $f_0$ satisfies
 	\begin{equation}
@@
 \end{theorem}
-Although~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} involves taking the maximum of several different terms, the important takeaway of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} is that if $n^{-1/2} \lesssim M \lesssim n^{1/d}$, then $\varphi$ has small worst-case risk as long as $f_0$ is separated from $0$ by at least $M^2(M^2n)^{-4/(4 + d)}$.
+Although~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} involves taking the maximum of several different terms, the important takeaway of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} is that if $n^{-1/2} \lesssim M \lesssim n^{(4 - d)/4d}$, then $\varphi$ has small worst-case risk as long as $f_0$ is separated from $0$ by at least $M^2(M^2n)^{-4/(4 + d)}$.
 Note that unlike in the estimation setting---where we measured loss in $L^2(P_n)$ error---the separation in~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} is measured in $L^2(P)$ norm.
@@
 
-Some other remarks:
-\begin{itemize}
-	\item As mentioned previously, when $d \geq 4$ the first order Sobolev space $H^1(\mc{X})$ does not continuously embed into $\Leb^4(\mc{X})$, and we do not know the optimal rates for regression testing over $H^1(\mc{X},M)$.
+As mentioned previously, when $d \geq 4$ the first order Sobolev space $H^1(\mc{X})$ does not continuously embed into $\Leb^4(\mc{X})$, and we do not know the optimal rates for regression testing over $H^1(\mc{X};M)$.
 On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then the Laplacian eigenmaps test with $K = n$, has small type II error whenever $\|f_0\|_P^2 \gtrsim Mn^{-1/2}$.
@@
 2021)} for details.
-	\item As in the estimation setting, the range of Sobolev ball radii $n^{-1/2} \lesssim M \lesssim n^{1/d}$ for which Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} implies that $\varphi$ is a rate-optimal test covers all those cases for which the critical radius $\epsilon(H^1(\mc{X};M))$ is both $\Omega(1/n)$ and $O(1)$.
-\end{itemize}
 
@@
 	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H_0^s(\mc{X},M)$ and $p \in C^{s - 1}(\mc{X})$.
-There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_ho}, then
+There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_ho}, then
 	\begin{equation}
@@
 	\label{asmp:parameters_testing_ho}
-	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy~\eqref{eqn:radius_ho}. Additionally,
+	The graph radius $\varepsilon$ satisfies~\eqref{eqn:radius_ho}, and the number of eigenvectors
 	\begin{equation}
 	\label{eqn:eigenvector_testing_ho}
-	K = \min\Bigl\{\floor{(M^2n)^{2d/(4 + d)}} \vee 1, n\Bigr\}.
+	K = \min\Bigl\{\floor{(M^2n)^{2d/(4s + d)}} \vee 1, n\Bigr\}.
 	\end{equation}
@@
 Suppose additionally $f_0 \in H_0^s(\mc{X},M)$, that $p \in C^{s-1}(\mc{X})$, and that $d \leq 4$.
-Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n \geq N$: if the Laplacian eigenmaps test $\varphi$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_ho}, and if $f_0$ satisfies
+Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n \geq N$: if the Laplacian eigenmaps test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_ho}, and if $f_0$ satisfies
 	\begin{equation}
@@
 \end{theorem}
-Similarly to the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that $\varphi$ is a minimax optimal test over $H_0^s(\mc{X})$ when $n^{-1/2} \lesssim M^2 \lesssim n^{1/d}$.
+Similarly to the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that $\varphi$ is a minimax optimal test over $H_0^s(\mc{X})$ when $n^{-1/2} \lesssim M \lesssim n^{(4s - d)/4d}$.
 However, unlike the first-order case, when $4 < d < 4s$ the minimax testing rate over $H_0^s(\mc{X})$ is still on the order of $M^2(M^2n)^{-4s/(4s + d)}$.
@@
 	Under the same setup as Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, but with $4 < d < 4s$.
-If the Laplacian eigenmaps test $\varphi$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, number of eigenvectors $K$ satisfying~\eqref{eqn:eigenvector_testing_ho}, and $\varepsilon = (M^2n)^{-1/(2(s - 1) + d)}$, and if
+If the Laplacian eigenmaps test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, number of eigenvectors $K$ satisfying~\eqref{eqn:eigenvector_testing_ho}, and $\varepsilon = (M^2n)^{-1/(2(s - 1) + d)}$, and if
 	\begin{equation}
@@
 Naively applying these results, one can show that $\wh{f}$ to $\wt{f}$, but only at a rate far slower than the optimal rates for regression.
-Of course when the index $K$ increases with $n$, as is necessary to optimally balance bias and variance, the issue only gets worse.
-Clearly, as a method for regression, the rate of convergence of Laplacian eigenmaps is much better than the rate implied by (what is currently known about) the concentration of individual eigenvectors around their continuum limits.
+Of course when the number of eigenvectors $K$ increases with $n$, as is necessary to optimally balance bias and variance of Laplacian eigenmaps, the issue only gets worse.
+Clearly, as a method for estimation and testing, the rate of convergence of Laplacian eigenmaps is much better than the rate implied by (what is currently known about) the concentration of individual eigenvectors around their continuum limits.
 
@@
 \label{eqn:estimation_biasvariance}
-\|\wh{f} - f_0\|_n^2 \leq \underbrace{\frac{\dotp{L^s f_0}{f_0}_n}{\lambda_{K}^s}}_{\textrm{bias}} + \underbrace{\frac{5K}{n} \vphantom{\frac{\dotp{L^s f_0}{f_0}_n}{\lambda_{K}^s}}}_{\textrm{variance}}.
+\|\wh{f} - f_0\|_n^2 \leq \underbrace{\frac{\dotp{L_{n,\varepsilon}^s f_0}{f_0}_n}{\lambda_{K}^s}}_{\textrm{bias}} + \underbrace{\frac{5K}{n} \vphantom{\frac{\dotp{L^s f_0}{f_0}_n}{\lambda_{K}^s}}}_{\textrm{variance}}.
 \end{equation}
@@
 \label{eqn:testing_biasvariance}
-\|f_0\|_n^2 \geq  \underbrace{\frac{\dotp{L^s f_0}{f_0}_n}{\lambda_{K}^s}}_{\textrm{bias}} + \underbrace{32\frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]}_{\textrm{variance}}.
-\end{equation}
-The quadratic form $\dotp{L_{n,\varepsilon}^s f_0}{f_0}_n$, eigenvalue $\lambda_{K}$, and empirical squared norm $\|f_0\|_n^2$ are each random variables that depend the random design points $X_1,\ldots,X_n$.
+\|f_0\|_n^2 \geq  \underbrace{\frac{\dotp{L_{n,\varepsilon}^s f_0}{f_0}_n}{\lambda_{K}^s}}_{\textrm{bias}} + \underbrace{32\frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]}_{\textrm{variance}}.
+\end{equation}
+The quadratic form $\dotp{L_{n,\varepsilon}^s f_0}{f_0}_n$, eigenvalue $\lambda_{K}$, and empirical squared norm $\|f_0\|_n^2$ are each random variables that depend on the random design points $X_1,\ldots,X_n$.
 We proceed to establish suitable upper and lower bounds on these quantities.
 
-\paragraph{Estimates on graph quadratic forms.}
+\paragraph{Estimates of graph quadratic forms.}
 In Proposition~\ref{prop:graph_seminorm_fo} we restate an upper bound on the Dirichlet energy $\dotp{L_{n,\varepsilon}f}{f}_n$ from \textcolor{blue}{Green et al. 2021}. 
@@
 \end{proposition}
-Proposition~\ref{prop:graph_seminorm_fo} follows by upper bounding the expectation of $\dotp{L_{n,\varepsilon}f}{f}_n$, which is the Dirichlet energy  $E_{P,\varepsilon}(f) := \dotp{L_{P,\varepsilon}f}{f}_P$, by (a constant times) the squared Sobolev norm $\|f\|_{H^1(\mc{X})}^2$.
+Proposition~\ref{prop:graph_seminorm_fo} follows by upper bounding the expectation of $\Ebb\dotp{L_{n,\varepsilon}f}{f}_n = \dotp{L_{P,\varepsilon}f}{f}_P$---where $L_{P,\varepsilon}$ is the non-local Laplacian operator defined in~\eqref{eqn:nonlocal_laplacian}---by (a constant times) the squared Sobolev norm $\|f\|_{H^1(\mc{X})}^2$, and an application of Markov's inequality.
 
@@
 \end{proposition}
-We now summarize the techniques used to prove Proposition~\ref{prop:graph_seminorm_ho}, which will help explain what role the conditions on $f_0$,$p$ and $\varepsilon$ play.
+We now summarize the techniques used to prove Proposition~\ref{prop:graph_seminorm_ho}, which will help explain what role the conditions on $f_0$, $p$ and $\varepsilon$ play.
 To upper bound $\dotp{L_{n,\varepsilon}^sf}{f}_n$ in terms of $\|f\|_{H^s(\mc{X})}^2$, we introduce an intermediate quantity: the \emph{order-s non-local Sobolev seminorm} $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$.
-This seminorm is defined with respect to $L_{P,\varepsilon}$, which is a non-local approximation to $\Delta_P$,
+This seminorm is defined with respect to the iterated non-local Laplacian $L_{P,\varepsilon}^s = L_{P,\varepsilon} \circ \cdots \circ L_{P,\varepsilon}$, where $L_{P,\varepsilon}$ is a non-local approximation to $\Delta_P$,
 \begin{equation}
@@
 We show that these pure bias terms are negligible when $\varepsilon = \omega(n^{-1/(2(s - 1) + d)})$.
-	\item For $x$ sufficiently in the interior of $\mc{X}$, we show that $L_{P,\varepsilon}^jf(x) \to \sigma_{\eta}^j \Delta_P^jf(x)$ as $\varepsilon \to 0$.
+	\item For $x$ sufficiently far from the boundary of $\mc{X}$---precisely $x \in \mc{X}$ such that $B(x,j\varepsilon) \subseteq \mc{X}$---we show that $L_{P,\varepsilon}^jf(x) \to \sigma_{\eta}^j \Delta_P^jf(x)$ as $\varepsilon \to 0$.
 Here $j = (s - 1)/2$ when $s$ is odd and $j = (s - 2)/2$ when $s$ is even.
 This step bears some resemblance to the analysis of the bias term in kernel smoothing, and requires that $p \in C^{s-1}(\mc{X})$.
-	\item On the other hand for $x$ sufficiently near the exterior of $\mc{X}$, $L_{P,\varepsilon}^jf(x)$ does not converge to $\Delta_P^jf(x)$.
+	\item On the other hand for $x$ sufficiently near the boundary of $\mc{X}$, $L_{P,\varepsilon}^jf(x)$ does not in general converge to $\sigma_{\eta}^j\Delta_P^jf(x)$.
 Instead, we use the zero-trace property of $f$ to show that $L_{P,\varepsilon}^jf(x)$ is small.
@@
 Together, these steps establish Proposition~\ref{prop:graph_seminorm_ho}.
-It is worth pointing out that we do not try to show $L_{P,\varepsilon}^sf(x) \to \Delta_{P}^sf(x)$.
-This may seem like a natural first step towards a simple proof that $\dotp{L_{P,\varepsilon}^sf}{f}_{P} \to \dotp{\Delta_P^sf}{f}_{P}$.
-The problem is that $\Delta_P^s$ is an order-$2s$ differential operator, whereas we assume that $f$ has only $s$ bounded derivatives.
-Instead we go for the slightly more complicated approach outlined above.
-
-% AG: I think this following might just be garbage, but I will leave it in in case it piques Ryan or Siva's interest.
-
-% Proposition~\ref{prop:graph_seminorm_ho}, and its proof, show that the iterated graph Laplacian $L_{n,\varepsilon}^s$ does a reasonable job of approximating a higher-order differential operator.
-The conditions we require --- that $\varepsilon$ be sufficiently large, that $f$ be zero-trace, and that $p$ have regularity of order $s - 1$ --- indicate that the iterated graph Laplacian is an imperfect, if adequate, tool for this job.
-This is no fault of the graph Laplacian.
-Rather, as pointed out by \textcolor{red}{(Sadhanala et al.
-2017)} its reflects that graph Laplacians were designed for a very general circumstance in which the only notions of derivation are first-differentials, divergences, and compositions of the two.
-Indeed, viewed in this light it is a pleasant surprise that iterated graph Laplacians approximate higher-order differentials at all.
-When one assumes Euclidean structure, as we do in this paper, there exist many methods (e.g.
-local linear embeddings, Hessian local linear embedding, local tangent space alignment) which leverage this structure to approximate differential operators in a more sophisticated manner.
-However, thus far little theoretical investigation has been done into even the pointwise behavior of these approaches.%
-
-\paragraph{Neighborhood graph eigenvalue.}
+It is worth pointing out that we do not try to establish the pointwise estimate $L_{P,\varepsilon}^sf \to \sigma_{\eta}^s\Delta_{P}^sf$ in $L^2(P)$.
+If we had such an estimate, it would immediately follow that $\dotp{L_{P,\varepsilon}^sf}{f}_{P} \to \sigma_{\eta}^s\dotp{\Delta_P^sf}{f}_{P}$.
+Unfortunately, we assume only that $f$ has $s$ bounded derivatives, while $\Delta_P^s$ is an order-$2s$ differential operator; thus in general $L_{P,\varepsilon}^sf$ may not approach $\sigma_{\eta}^s\Delta_{P}^sf$ as $\varepsilon \to 0$.
+Instead we opt for the slightly more complicated approach outlined above, in which we only ever need show that $L_{P,\varepsilon}^jf(x) \to \sigma_{\eta}^j \Delta_P^jf(x)$ for some $j < s/2$.
+
+\paragraph{Neighborhood graph eigenvalues.}
 On the other hand, several recent works \citep{burago2014,garciatrillos18,calder2019} have analyzed the convergence of $\lambda_{k}$ towards $\lambda_{k}(\Delta_P)$.
@@
 	\end{equation}
-	then with probability at least $1 - \exp\{-(cn \wedge 1/\delta)\}$.
+	then with probability at least $1 - \exp\{-(cn \wedge 1/\delta)\}$,
 	\begin{equation}
@@
 \end{proposition}
-To prove Proposition~\ref{prop:empirical_norm_sobolev}, we use the Gagliardo-Nirenberg interpolation inequality to control the $4$th moment of $f$ in terms of $\|f\|_P$ and $|f|_{H^s(\mc{X})}$, then invoke a one-sided Bernstein's inequality as in \cite[Section 14.2]{wainwright2019}.
+To prove Proposition~\ref{prop:empirical_norm_sobolev}, we use the Gagliardo-Nirenberg interpolation inequality (see e.g.
+Theorem~(?) of~\citep{evans10}) to control the $4$th moment of $f$ in terms of $\|f\|_P$ and $|f|_{H^s(\mc{X})}$, then invoke a one-sided Bernstein's inequality as in \cite[Section 14.2]{wainwright2019}.
 Note carefully that the statement~\eqref{eqn:empirical_norm_sobolev} is \emph{not} a uniform guarantee over all $f \in H^s(\mc{X};M)$, as such a statement cannot hold in the sub-critical regime ($2s \leq d$).
 Fortunately, a pointwise bound---meaning a bound that holds with high probability for a single $f \in H^s(\mc{X})$---is sufficient for our purposes.
+% (AG) Ask Ryan for reference to Evans.
+% (SB via AG): More discussion about the uniform guarantee point.
 
@@
 In fact the aforementioned edge sparsification algorithms are overkill for our needs.
-For one thing, they are designed to work when $\sigma$ is much larger than $1$, whereas in order for $\wc{f}$ to be rate-optimal setting $\sigma$ to be any constant greater than $1$, say $\sigma = 2$, is sufficient.
+For one thing, they are designed to work when $\sigma$ is very close to $1$, whereas in order for $\wc{f}$ to be rate-optimal setting $\sigma$ to be any constant greater than $1$, say $\sigma = 2$, is sufficient.
 Additionally, edge sparsification algorithms are traditionally designed to work in the worst-case, where no assumptions are made on the structure of the graph $G$.
 But the geometric graphs we consider in this paper exhibit a special structure, in which very roughly speaking no single edge is a bottleneck.
-As pointed out by~\citet{sadhanala16b}, in this special case there exist far simpler and faster methods for sparsification, which at least empirically seem to do the job.
+As pointed out by~\citet{sadhanala16b}, in this special case there are far simpler and faster methods for sparsification, which at least empirically seem to do the job.
 
@@
 	\label{eqn:radius_estimation_manifold}
-	C_0\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/m}, n^{-1/(2(s - 1) + m)}\biggr\} \leq \varepsilon \leq \min\{i_0, K^{-1/m}\}
-	\end{equation}
-	and
+	C_0\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/m}, n^{-1/(2(s - 1) + m)}\biggr\} \leq \varepsilon \leq \min\{i_0, K^{-1/m}\},
+	\end{equation}
+	where we recall that $i_0$ is a lower bound on the injectivity radius of $\mc{X}$. Additionally,
 	\begin{equation*}
@@
 	Suppose Model~\ref{def:model_manifold}, and additionally $f_0 \in H^s(\mc{X},M)$ and $p \in C^{s - 1}(\mc{X})$ for $s \leq 3$.
-There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with kernel $\eta$ satisfying~\ref{asmp:kernel_manifold}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_manifold}, then
+There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_manifold}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_manifold}, then
 	\begin{equation}
@@
 	\label{asmp:parameters_testing_manifold}
-	The graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy~\eqref{eqn:radius_estimation_manifold}. Additionally, the
+	The graph radius $\varepsilon$ satisfies~\eqref{eqn:radius_estimation_manifold}, and the number of eigenvectors
 	\begin{equation*}
@@
 Suppose additionally $f_0 \in H^s(\mc{X},M)$, that $p \in C^{s-1}(\mc{X})$, and that $s \leq 3$ and $m \leq 4$.
-Then there exist constants $c$, $C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed kernel $\eta$ satisfying~\ref{asmp:kernel_manifold}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_manifold}, and if $f_0$ satisfies
+Then there exist constants $c$, $C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_manifold}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_manifold}, and if $f_0$ satisfies
 	\begin{equation}
 	\label{eqn:laplacian_eigenmaps_testing_criticalradius_manifold}
-	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-4s/(4s + m) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{2s/d}} \biggr) \vee \frac{1}{n},
+	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-4s/(4s + m) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{2s/m}} \biggr) \vee \frac{1}{n},
 	\end{equation}
@@
 The difference is that when $\mc{X}$ is a manifold with intrinsic dimension $m$, we can prove analogous results to Propositions~\ref{prop:graph_seminorm_fo}-\ref{prop:graph_eigenvalue}, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
-	\item Unlike in the full-dimensional case, our upper bounds on the estimation and testing error of Laplacian eigenmaps match the minimax rate only when $s \leq 3$.
-Our upper bounds when $s \geq 4$ follow from the embedding $H^s(\mc{X};M) \subset H^{3}(\mc{X};M)$, i.e they match the rates we get just by assuming that the $3$rd order derivative is bounded, and are clearly suboptimal.
-
-	We now explain this discrepancy.
-At a high level, thinking of the graph $G$ as an estimate of the manifold $\mc{X}$, we incur some error by using Euclidean distance rather than geodesic distance to form the edges of $G$.
+	\item Unlike in the full-dimensional case, in the manifold setting our upper bounds on the estimation and testing error of Laplacian eigenmaps match the minimax rate only when $s \leq 3$.
+When $s \geq 4$, the containment $H^s(\mc{X};1) \subset H^{3}(\mc{X};1)$ (taking the radius of the Sobolev ball $M = 1$ for simplicity) implies that the Laplacian eigenmaps estimator $\wh{f}$ has in-sample mean-squared error of at most $n^{-6/(6 + m)}$, and that the Laplacian eigenmaps test has small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-12/(12 + d)}$; however, these are slower than the usual minimax rates.
+
+	We now explain this discrepancy, between the full-dimensional and manifold settings.
+At a high level, thinking of the graph $G$ as an estimate of the manifold $\mc{X}$, we incur some error in this estimate by using Euclidean distance rather than geodesic distance to form the edges of $G$.
 This is in contrast with the full-dimensional setting, where the Euclidean metric exactly coincides with the geodesic distance for all points $x,z \in \mc{X}$ that are sufficiently close to each other and far from the boundary of $\mc{X}$.
@@
 
-	\item Finally, when $m > 4$, we get an upper bound on testing error equivalent to that of Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}, except with the ambient dimension $d$ replaced by intrinsic dimension $m$.
+	\item Finally, as in the full-dimensional case, when the intrinsic dimension $m > 4$ we cannot choose the graph radius $\varepsilon$ and number of eigenvectors $K$ to optimally balance bias and variance.
+Instead, reasoning as in the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} shows that when $1 \leq s \leq 3$, the Laplacian eigenmaps test has critical radius as given by~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_ho_suboptimal}, but with the ambient dimension $d$ replaced by $m$.
 \end{itemize}
+% (SB via AG): Maybe some comment on the difference between testing and estimation.
 
@@
 
-Finally, using a Gagliardo-Nirenberg inequality for functions on compact Riemmanian manifolds, we obtain a lower bound on empirical norm $\|f\|_n$ under the hypotheses of Model~\ref{def:model_manifold}.
+Finally, we have the following lower bound on the empirical norm $\|f\|_n$ under the hypotheses of Model~\ref{def:model_manifold}. 
 \begin{proposition}
@@
 \end{proposition}
+We prove Proposition~\ref{prop:empirical_norm_sobolev_manifold} in a parallel manner to its flat Euclidean counterpart (Proposition~\ref{prop:empirical_norm_sobolev}), by using a Gagliardo-Nirenberg inequality for functions on compact Riemannian manifolds, and then applying a one-sided Bernstein's inequality.
+Finally, combining Propositions~\ref{prop:graph_seminorm_manifold}-\ref{prop:empirical_norm_sobolev_manifold} with the conditional-on-design bias-variance decompositions~\eqref{eqn:estimation_biasvariance} and \eqref{eqn:testing_biasvariance} leads to the conclusions of Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold}.
 
@@
 We now formally define our approach to extension by kernel smoothing.
-For a kernel function $\psi(\cdot): [0,\infty) \to (-\infty,+\infty)$, bandwidth $h > 0$, and a distribution $Q$, the \emph{Nadaraya-Watson kernel smoother} $T_{h,Q}$ is given by
+For a kernel function $\psi(\cdot): [0,\infty) \to (-\infty,+\infty)$, bandwidth $h > 0$, and a distribution $Q$, the \emph{Nadaraya-Watson kernel smoother} $T_{Q,h}$ is given by
 \begin{equation*}
@@
 where $d_{Q,h}(x) := \int_{\Omega} \psi\bigl(\|z - x\|/\varepsilon\bigr) \,dQ(z)$.
-For convenience, we will write $T_{\varepsilon,n}f(x) := T_{\varepsilon,P_n}f(x)$, and $d_{n,h}(x) := n \cdot d_{P_n,h}(x)$.
-We extend the Laplacian eigenmaps estimator by passing the kernel smoother $T_{h,n}$ over it, that is we consider the estimator $T_{h,n}\wh{f}$, which is defined at every $x \in \mc{X}$ (indeed, at every $x \in \Rd$).
-Note that ``extension'' here is a slight abuse of nomenclature, since $T_{h,n}\wh{f}(X_i)$ and $\wh{f}_i$ may not agree in-sample.
+For convenience, we will write $T_{n,h}f(x) := T_{P_n,h}f(x)$, and $d_{n,h}(x) := n \cdot d_{P_n,h}(x)$.
+We extend the Laplacian eigenmaps estimator by passing the kernel smoother $T_{n,h}$ over it, that is we consider the estimator $T_{n,h}\wh{f}$, which is defined at every $x \in \mc{X}$ (indeed, at every $x \in \Rd$).
+Note that ``extension'' here is a slight abuse of nomenclature, since $T_{n,h}\wh{f}(X_i)$ and $\wh{f}_i$ may not agree in-sample.
 
@@
 
-In Lemma~\ref{lem:kernel_smoothing_insample}, we consider an arbitrary estimator $\wc{f} \in L^2(P_n)$.
-We show that the out-of-sample error $\|T_{n,h}\wc{f} - f_0\|_P^2$ can be upper bounded by three terms--- (a constant times) the in-sample error $\|\wc{f} - f_0\|_n^2$, and variance and bias terms that arise naturally in the analysis of kernel smoothing over noiseless data..
+In Lemma~\ref{lem:kernel_smoothing_insample}, we consider an arbitrary estimator $\wc{f} \in \Reals^n$.
+We show that the out-of-sample error $\|T_{n,h}\wc{f} - f_0\|_P^2$ can be upper bounded by three terms--- (a constant times) the in-sample error $\|\wc{f} - f_0\|_n^2$, and variance and bias terms that arise naturally in the analysis of kernel smoothing over noiseless data.
 We shall assume the following conditions on $\psi$ and $h$.
@@
 	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $\wc{f} \in L^2(P_n)$, $f_0 \in H^1(\mc{X})$ and $p \in C^1(\mc{X})$.
-If the kernel smoothing estimator $T_{h,n}\wc{f}$ is computed with kernel $\psi$ satisfying~\ref{asmp:kernel} and bandwidth $h$ satisfying~\ref{asmp:bandwidth}, it holds that
+If the kernel smoothing estimator $T_{n,h}\wc{f}$ is computed with a kernel $\psi$ satisfying~\ref{asmp:kernel} and bandwidth $h$ satisfying~\ref{asmp:bandwidth}, it holds that
 	\begin{equation}
 	\label{eqn:kernel_smoothing_insample}
-	\|T_{n,h}\wc{f} - f_0\|_P^2 \leq C\biggl(\|\wc{f} - f_0\|_n^2 + \frac{1}{\delta} \cdot \frac{h^2}{nh^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{h,P}f_0 - f_0\|_P^2\biggr),
+	\|T_{n,h}\wc{f} - f_0\|_P^2 \leq C\biggl(\|\wc{f} - f_0\|_n^2 + \frac{1}{\delta} \cdot \frac{h^2}{nh^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{P,h}f_0 - f_0\|_P^2\biggr),
 	\end{equation}
@@
 On the other hand the bias term is typical.
-When $\psi$ is an order-$s$ kernel, a standard analysis shows that the $\|T_{h,P}f_0 - f_0\|_P^2 \lesssim \varepsilon^{2s}$.
+When $\psi$ is an order-$s$ kernel, a standard analysis shows that the $\|T_{P,h}f_0 - f_0\|_P^2 \lesssim \varepsilon^{2s}$.
 \begin{enumerate}[label=(K\arabic*)]
@@
 \end{enumerate}
-
-\textcolor{red}{(TODO): I have worked out the math for this result, but I am a little surprised at the requirement that $\psi$ need annihilate polynomials up to degree $s + d - 2$.
-Will need to double check closely, since I could not find a reference analyzing kernel smoothing using a radially symmetric higher order kernel when $d > 1$.}
+% (SB via AG): Give an example of a kernel which satisfies these requirements.
 
@@
 \end{equation}
-\eqref{eqn:kernel_smoothing_insample2} tells us that the additional error incurred by passing a kernel smoother over an in-sample estimator $\wc{f}$ is negligible compared to the minimax rate of estimation.
-Consequently, if $\wc{f}$ converges at the minimax rate in $L^2(P_n)$, then $T_{n,h}\wc{f}$ will converge at the minimax rate in $L^2(P)$.
-It follows immediately from Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} (when $s = 1$) or Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} (when $s > 1$), that $T_{h,n}\wh{f}$ achieves the optimal rate of convergence in $L^2(P)$.
+This analysis tells us that the additional error incurred by passing a kernel smoother over an in-sample estimator $\wc{f}$ is negligible compared to the minimax rate of estimation.
+Consequently, if $\wc{f}$ converges at the minimax rate in in-sample mean squared error, then $T_{n,h}\wc{f}$ will converge at the minimax rate in $L^2(P)$.
+It follows immediately from Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} (when $s = 1$) or Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} (when $s > 1$), that $T_{n,h}\wh{f}$ achieves the optimal rate of convergence in $L^2(P)$.
 
@@
 	\begin{itemize}
-		\item If $f_0 \in H^1(\mc{X};M)$, the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_fo}, and the out-of-sample extension $T_{h,n}\wh{f}$ is computed with bandwidth $h = n^{-1/d}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel}, then
-		\begin{equation*}
-		\|T_{h,n}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2s/(2s + d)}.
-		\end{equation*}
-		\item If $f_0 \in H_0^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$ for some $s \in \mathbb{N}, s > 1$, and the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_ho}, and the out-of-sample extension $T_{h,n}\wh{f}$ is computed with bandwidth $h = n^{-1/(2(s - 1) + d)}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel} and~\ref{asmp:ho_kernel}, then
-		\begin{equation*}
-		\|T_{h,n}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2s/(2s + d)}.
+		\item If $f_0 \in H^1(\mc{X};M)$, the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_fo}, and the out-of-sample extension $T_{n,h}\wh{f}$ is computed with bandwidth $h = n^{-1/d}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel}, then
+		\begin{equation*}
+		\|T_{n,h}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2s/(2s + d)}.
+		\end{equation*}
+		\item If $f_0 \in H_0^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$ for some $s \in \mathbb{N}, s > 1$, and the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_ho}, and the out-of-sample extension $T_{n,h}\wh{f}$ is computed with bandwidth $h = n^{-1/(2(s - 1) + d)}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel} and~\ref{asmp:ho_kernel}, then
+		\begin{equation*}
+		\|T_{n,h}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2s/(2s + d)}.
 		\end{equation*}
@@
 	\item Since $T_{n,h}\wh{f}$ is defined out-of-sample, we can use sample splitting or cross validation methods to tune hyperparameters, which we could not do for the original estimator $\wh{f}$.
-For instance, we can (i) split the sample into two halves, (ii) use the first half to compute $T_{h,n}\wh{f}$ for various values of $\varepsilon$, $h$, and $K$, (iii) choose the optimal values of these three hyperparameters by minimizing error on the held out set.
+For instance, we can (i) split the sample into two halves, (ii) use the first half to compute $T_{n,h}\wh{f}$ for various values of $\varepsilon$, $h$, and $K$, (iii) choose the optimal values of these three hyperparameters by minimizing error on the held out set.
 Practically speaking, cross-validation is one of the most common approaches to choosing hyperparameters.
@@
 In other words, it leads to estimators that are rate-optimal (up to $\log n$ factors), even when $s$ is unknown.
-Similar arguments should imply that $T_{h,n}\wh{f}$ is adaptive in this sense when $\varepsilon,h$ and $K$ are chosen by sample splitting.
+Similar arguments should imply that $T_{n,h}\wh{f}$ is adaptive in this sense when $\varepsilon,h$ and $K$ are chosen by sample splitting.
 \textcolor{red}{(TODO): Siva would like me to fill in some details?}
-	\item There exist many other approaches to extending a function $f$ using only evaluations $\{f(X_1),\ldots,f(X_n)\}$.
+	\item There exist other approaches to extending a function $f$ from its $\{f(X_1),\ldots,f(X_n)\}$: for instance, minimum-norm interpolation in some RKHS or Banach space \textcolor{blue}{(Rieger2008, Belkin2018)}, or 1-nearest neighbors regression \textcolor{blue}{(citation)}.
 We consider extension by kernel smoothing because it is a simple and statistically optimal procedure that does not require any knowledge of the domain $\mc{X}$ or distribution $P$---as we have argued, this latter property is one of the main selling points of Laplacian eigenmaps as a tool for nonparametric regression.
-
-	We now comment on a few alternative methods. One such approach is minimum norm interpolation. Here one defines a normed space $(\mc{F},\|\cdot\|_{\mc{F}})$ and then solves the optimization problem
-	\begin{equation*}
-	\min_{u \in \mc{F}} \|u\|_{\mc{F}},~~\textrm{such that}~~ u(X_i) =  f(X_i)~~\textrm{for $i = 1,\ldots,n$.}
-	\end{equation*}
-	A particularly popular version of this general approach takes $\mc{F}$ to be an RKHS \textcolor{red}{(Rieger2008, Belkin2018)}, which encompasses thin-plate spline interpolation (where $\mc{F} = H^s(\mc{X})$ for $s > d/2$) as a special case.
-Naturally, this approach works well when $f$ is close to a function $u \in \mc{F}$ with reasonably small norm.
-This holds true when $f \in H^s(\mc{X};M)$ and $s > d/2$, but as already discussed when $s \leq d/2$ the Sobolev space $H^s(\mc{X})$ is not an RKHS, and in fact when $s \leq d/2$ thin-plate spline interpolation is ill-posed \citep{green93}.
-Another, arguably simpler approach is to extend $f$ to be piecewise constant on the Voronoi tessellation induced by $X_1,\ldots,X_n$, or equivalently to perform $1$-nearest neighbors regression on $f$.
-However, this approach is theoretically optimal only when $f_0$ is Lipschitz, in contrast to the kernel smoothing method we propose and study.
-
-	None of these three methods are intrinsically linked to Laplacian eigenmaps.
-This is in one sense a strength, since they can be used to extrapolate any estimator defined only in-sample.
-But it is also potentially a weakness.
-Each of these methods have their own approximation and estimation errors (bias and variance) which can be fundamentally different than those of Laplacian eigenmaps, and there is a danger that in extrapolating the Laplacian eigenmaps estimator in this way we are taking the ``worst of both worlds''.
-Our theory shows that when the data model is Model~\ref{def:model_flat_euclidean} and we perform extrapolation by kernel smoothing, this is not a problem, at least in a minimax sense.
+One potential drawback to kernel smoothing is that it can change the value of the estimate at data, meaning $\wh{f} \neq (\wc{f}(X_1),\ldots, \wc{f}(X_n))$.
+More sophisticated procedures such as Nystr\"{o}m extension of the eigenvectors $v_1,\ldots,v_K$ \textcolor{blue}{(citation)} inherit the generality of kernel smoothing while being genuine interpolators; however, their properties are substantially more difficult to analyze.
 
@@
 
+% (SB via AG): Change the name of this to ``spectral projection'' or something like that.
 \paragraph{Nonparametric least squares.}
@@
 
+% (SB via AG): Maybe also worth discussing the ill-posed counterpart?
+
 \paragraph{Kernel smoothing.}
-It is also natural to ask whether the two stage estimator $T_{h,n}\wh{f}$ defined in Section~\ref{sec:out_of_sample} has any advantage over the simpler approach of directly kernel smoothing the responses, i.e.
-using the estimator $T_{h,n}Y$ (possibly for a different choice of $h$).
-In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{h,n}\wh{f}\|_P^2$ is of a strictly lower order than $\inf_{h'} \Ebb\|f_0 - T_{h',n}Y\|_P^2$.
+It is also natural to ask whether the two stage estimator $T_{n,h}\wh{f}$ defined in Section~\ref{sec:out_of_sample} has any advantage over the simpler approach of directly kernel smoothing the responses, i.e.
+using the estimator $T_{n,h}Y$ (possibly for a different choice of $h$).
+In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{n,h}\wh{f}\|_P^2$ is of a strictly lower order than $\inf_{h'} \Ebb\|f_0 - T_{h',n}Y\|_P^2$.
 This is possible because Laplacian eigenmaps induces a completely different bias than kernel smoothing.
@@
 
+% (SB via AG): Add references, and clarify what is meant by used to.
 Much is known about classical spectral projection methods beyond their rate optimality.
@@
 
+% AG: I think this following might just be garbage, but I will leave it in in case it piques Ryan or Siva's interest.
+
+% Proposition~\ref{prop:graph_seminorm_ho}, and its proof, show that the iterated graph Laplacian $L_{n,\varepsilon}^s$ does a reasonable job of approximating a higher-order differential operator.
+The conditions we require --- that $\varepsilon$ be sufficiently large, that $f$ be zero-trace, and that $p$ have regularity of order $s - 1$ --- indicate that the iterated graph Laplacian is an imperfect, if adequate, tool for this job.
+This is no fault of the graph Laplacian.
+Rather, as pointed out by \textcolor{red}{(Sadhanala et al.
+2017)} its reflects that graph Laplacians were designed for a very general circumstance in which the only notions of derivation are first-differentials, divergences, and compositions of the two.
+Indeed, viewed in this light it is a pleasant surprise that iterated graph Laplacians approximate higher-order differentials at all.
+When one assumes Euclidean structure, as we do in this paper, there exist many methods (e.g.
+local linear embeddings, Hessian local linear embedding, local tangent space alignment) which leverage this structure to approximate differential operators in a more sophisticated manner.
+However, thus far little theoretical investigation has been done into even the pointwise behavior of these approaches.%
+
 \bibliographystyle{plainnat}

### 2021-08-13 manuscripts/laplacian_eigenmaps/sections/introduction.tex
Fixed all of Siva's small changes. Comprehensive changes made to introduction of Laplacian Eigenmaps paper.

\section{Introduction}
\label{sec:introduction}
Two fundamental goals of statistics, data science, and machine learning are to learn and leverage structure of an unknown distribution, on the basis of samples drawn from said distribution. Neighborhood graph Laplacians, and their associated spectra, have proven to be powerful and general tools for both purposes. They have been successfully used for various statistical tasks such as (spectral) clustering, manifold learning, level-set estimation, semi-supervised learning, etc.

By this point there exists a rich literature (e.g. \citep{koltchinskii2000,belkin07,vonluxburg2008,burago2014,shi2015,singer2017,garciatrillos18,trillos2019, calder2019, cheng2021,dunson2021}) explaining this practical success from a theoretical perspective. Loosely speaking, these works justify graph spectral methods by showing that the eigenvectors and eigenvalues of graph Laplacians are empirical approximations of population-level objects: the eigenvalues and eigenfunctions of a density-weighted Laplace-Beltrami operator. These population-level objects in turn characterize various structural aspects of the density, such as its high- and low-density regions, the shape and intrinsic dimension of its support, and so forth. Together, these insights give a reasonably complete explanation of how graph spectra \emph{learn} the structure of an underlying distribution. In contrast, in this paper we develop theory showing that graph spectral methods can effectively \emph{leverage} structure to perform downstream (supervised) learning tasks. Our main goal is to show that spectral projection procedures for nonparametric estimation and testing---which use eigenvectors of a neighborhood graph Laplacian as features in linear regression---are statistically optimal.

The regression methods we consider are based on \emph{Laplacian eigenmaps}, first introduced by~\cite{belkin03a} \footnote{Technically speaking, as used by~\cite{belkin03a} ``Laplacian eigenmaps'' refers to embedding each design point $X_i$ according to coordinates defined by the first-$K$ graph Laplacian eigenvectors; that is, mapping $X_i \mapsto (v_{i,1},\ldots,v_{i,K})$. What we call Laplacian eigenmaps is a two-step supervised method: the first step is the embedding step, and the second step computes the estimate $\wh{f}$ using linear regression.  Since in this work we only ever deal with the supervised method, and since the choice of embedding step is the far more important of the two steps, we take the liberty of referring to the entire procedure as Laplacian eigenmaps}. Given samples $(X_1,Y_1),\ldots,(X_n,Y_n)$, Laplacian eigenmaps first constructs a neighborhood graph---with vertices at the design points $X_1,\ldots,X_n$ and weighted edges between pairs of sufficiently close points $X_i$ and $X_j$---and computes the corresponding (unweighted) graph Laplacian. The Laplacian eigenmaps estimate $\wh{f}$ is simply the projection of the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto the subspace spanned by a subset of the graph Laplacian eigenvectors. The empirical norm $\|\wh{f}\|_n^2$ can be used as a statistic to test whether any signal is present: that is, whether the regression function $f_0(x) = \mathbb{E}[Y|X = x]$ is anywhere non-zero. (The precise definition of these methods is given in Section~\ref{}.) We study these methods through the lens of a minimax analysis, and give upper bounds on the error of the Laplacian eigenmaps estimator and test which imply that they are minimax optimal over certain smoothness classes.

More specifically, we consider the worst-case risk of Laplacian eigenmaps over normed balls in Sobolev spaces. Roughly speaking, for a domain $\mc{X} \subseteq \Rd$, the order-$s$ (Hilbert-)Sobolev space $H^s(\mc{X})$ consists of those functions which are $s$-times weakly differentiable, with derivatives bounded in $L^2(\mc{X})$ norm. Under standard regularity conditions, the minimax estimation and testing rates for Sobolev spaces are well-known: when using mean-squared error as the loss, the estimation rate is $n^{-2s/(2s + d)}$, and the testing rate is $n^{-4s/(4s + d)}$.  These rates are achieved by various estimators, but classical spectral projection methods\footnote{We use ``classical'' to refer to spectral projection methods using eigenfunctions of a (fixed) Laplacian or Laplace-Beltrami operator, as distinct from Laplacian eigenmaps which uses eigenvectors of a (data-dependent) graph Laplacian matrix.}---which use the eigenfunctions $\psi_1,\psi_2,\ldots$ of a Laplace-Beltrami operator defined on $\mc{X}$---are particularly well-suited for regression over Sobolev spaces. This is because Sobolev spaces can be equivalently defined in a spectral manner: they consist of those functions $f$ whose Fourier coefficients $\theta_k := \int_{\mc{X}} f \psi_k$ are bounded in a particular sequence norm. For this reason, classical spectral projection methods have excellent statistical properties; see e.g.~\cite{tsybakov2008_book} and \cite{ingster2009} for analysis of estimation and testing risk in the special case where the design distribution is uniform over the unit cube, and Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} in Section~\ref{subsec:minimax_rates_sobolev} of our paper, which give equivalent upper bounds under more general conditions. However, these classical spectral methods require a priori knowledge of the design distribution, which is often not available.

On the other hand, Laplacian eigenmaps implicitly learns the structure of the design distribution, and serves as a data-dependent approximation to classical spectral projection that can be used even when the design distribution is unknown. Of course, this comes with a tradeoff. Intuitively, using an empirical approximation to the ``right'' population-level basis will incur some additional error; the question is whether this additional error is enough to meaningfully degrade the overall statistical properties.

We find that in many cases, the answer is no: under appropriate regularity conditions, when correctly tuned Laplacian eigenmaps is minimax optimal over Sobolev spaces. Specifically, our main contributions are as follows:
\begin{itemize}
	\item Over the Sobolev space $H^{s}(\mc{X})$, we establish the Laplacian eigenmaps estimator $\wh{f}$ has in-sample mean-square error on the order of $n^{-2s/(2s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d$. A test based on the statistic $\|\wh{f}\|_n^2$ has a critical radius on the order of $n^{-4s/(4s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d \in \{1,2,3,4\}$. 
	\item The estimator $\wh{f}$ is defined only in-sample. We propose a simple out-of-sample extension based on kernel smoothing, and show that this extension adds negligible error, resulting in an estimator which is defined on all of $\mc{X}$ and has out-of-sample mean-squared error also on the order of $n^{-2s/(2s + d)}$.
	\item We also consider the behavior of Laplacian eigenmaps when the data satisfies a \emph{manifold hypothesis}, meaning the design points $X_1,\ldots,X_n$ belong to some low-dimensional submanifold of $\Rd$. In this case, it is known that the minimax rates over Sobolev spaces depend on the intrinsic dimension $m$ of the manifold: explicitly, the estimation rate is $n^{-2s/(2s + m)}$ and the testing rate is $n^{-4s/(4s + m)}$. We give upper bounds implying that Laplacian eigenmaps achieves these rates so long as $s \in \{1,2,3\}$, and demonstrating that Laplacian eigenmaps is a~\emph{domain adaptive} estimator.
\end{itemize}
(In all cases, our bounds also depend optimally on the radius $M$ of the Sobolev ball under consideration.)

To derive these results, we develop an analysis framework reminiscent of that used in the analysis classical spectral projection estimators, but ultimately quite different. The chief difference is in the approximation error (bias). For our purposes we must establish that, with high probability, the eigenvectors of graph Laplacians effectively approximate continuum Sobolev functions, whereas in the classical case this follows immediately from the aforementioned spectral characterization of Sobolev spaces. To tighly upper bound the approximation error of Laplacian eigenmaps, we rely on several recent results regarding concentration of the spectra of graph Laplacians, and prove some additional results which may be of independent interest.

Our work establishes that Laplacian eigenmaps shares many of the optimality properties of classical spectral projection. However, for some values of $s$ (number of derivatives) and $d$ (dimension), there do exist gaps between our upper bounds on the error of Laplacian eigenmaps and the minimax rates. Although we do not give corresponding lower bounds verifying the tightness of our analysis, we believe these gaps reflect the true behavior of the method rather than some looseness in our analysis; we comment more on this at relevant parts in the text. For completeness, we summarize all of our upper bounds---those which match the minimax rates, and those which do not---in Tables~\ref{tbl:estimation_rates} and~\ref{tbl:testing_rates}.
\begin{table}
	\begin{center}
		\begin{tabular}{p{.2\textwidth} | p{.14\textwidth} p{.12\textwidth} }
			Smoothness order & Flat Euclidean (Model~\ref{def:model_flat_euclidean}) & Manifold (Model~\ref{def:model_manifold}) \\
			\hline
			$s \leq 3$ & $\bm{n^{-2s/(2s + d)}}$ & $\bm{n^{-2s/(2s + m)}}$ \\
			$s > 3$  & $\bm{n^{-2s/(2s + d)}}$ & $n^{-6/(6 + m)}$
		\end{tabular}
	\end{center}
	\caption{Summary of Laplacian eigenmaps estimation rates over Sobolev balls. Bold font marks minimax optimal rates. In each case, rates hold for all $d \in \mathbb{N}$ (under Model~\ref{def:model_flat_euclidean}), and for all $m \in \mathbb{N}, 1 < m < d$ (under Model~\ref{def:model_manifold}). Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps estimator is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $n^{-1/2} \lesssim M \lesssim n^{1/d}$.}
	\label{tbl:estimation_rates}
\end{table}

\begin{table}
	\begin{center}
		\begin{tabular}{p{.175\textwidth} p{.175\textwidth} | p{.14\textwidth} p{.12\textwidth} }
			Smoothness order & Dimension & Flat Euclidean (Model~\ref{def:model_flat_euclidean}) & Manifold (Model~\ref{def:model_manifold}) \\
			\hline
			\multirow{2}{*}{$s = 1$} & $\dim(\mc{X}) < 4$ & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
			& $\dim(\mc{X}) \geq 4$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
			\hline
			\multirow{3}{*}{$s = 2$ or $3$} & $\dim(\mc{X}) \leq 4$  & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
			& $4 <\dim(\mc{X}) < 4s$  & $n^{-2s/(2(s - 1) + d)}$ & $n^{-2s/(2(s - 1) + m)}$\\
			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
			\hline
			\multirow{3}{*}{$s > 3$} & $\dim(\mc{X}) \leq 4$ & $\bm{n^{-4s/(4s + d)}}$ & $n^{-12/(12 + d)}$ \\
			& $4 < \dim(\mc{X}) < 4s$ & $n^{-2s/(2(s - 1) + d)}$ & $n^{-6/(4 + m)}$ \\
			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
		\end{tabular}
	\end{center}
	\caption{Summary of Laplacian eigenmaps testing rates over Sobolev balls. Bold font marks minimax optimal rates. Rates when $d > 4s$ assume that $f_0 \in L^4(P,M)$. Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $n^{-1/2} \lesssim M \lesssim n^{1/d}$.}
	\label{tbl:testing_rates}
\end{table}

\paragraph{Related work.}

The majority of work on supervised learning using graphs adopts a \emph{fixed design} perspective, treating the design points $x_1,\ldots,x_n$ as vertices of a fixed graph, and carrying out inference with respect to the conditional mean vector $(f_0(x_1),\ldots,f_0(x_n))$. In this setting, tight upper bounds have been established that certify the optimality of graph-based methods for estimation \citep{wang2016,hutter2016,sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}) and testing \citep{sharpnack2010identifying,sharpnack2013b,sharpnack2013,sharpnack2015} over different ``function'' classes (in quotes because these classes really model the $n$-dimensional vector of evaluations). In particular,~\citet{sadhanala16} and~\citet{sharpnack2015} analyze the Laplacian eigenmaps estimator and test statistic, respectively. This setting is quite general, because the graph need not be a geometric graph defined on a vertex set which belongs to Euclidean space. On the other hand, depending on the data-collection process, it may be unnatural to model the design points as being a priori fixed, and the estimand as being a vector which exhibits a discrete notion of ``smoothness'' over this fixed design. Instead, we adopt the \emph{random design} perspective, and seek to estimate a function that we assume exhibits a more classical notion of smoothness.

In the random design setting the analysis is more challenging, and only a few works have studied the matter. \cite{zhou2011} provide an asymptotic analysis of the Laplacian eigenmaps estimator in the semi-supervised setting, as the number of unlabeled samples tends to infinity; in this case the estimator reduces to the classical spectral projection estimator, and there is no extra approximation error due to using data-dependent eigenvectors. \cite{lee2016} consider the \emph{diffusion maps} estimator---which uses the eigenvectors of a differently normalized graph Laplacian---in a similar supervised setting to our own, but prove suboptimal rates of convergence relative to the minimax rates. Finally, \citet{trillos2020,green2021} study a penalized least squares method for estimation, which uses a quadratic form involving the Laplacian as a penalty. This method is optimal only for the first-order Sobolev space $H^1(\mc{X})$, and even then only when $d \in \{1,2,3,4\}$, in contrast to the Laplacian eigenmaps estimator which is optimal for all $s$ and $d$; we discuss this striking contrast more in Section~\ref{sec:discussion}.

\paragraph{Roadmap.}
We now outline the structure of the rest of this paper. In Section~\ref{sec:setup_main_results}, we formally setup the models and methods we study. Propositions~\ref{prop:spectral_series_estimation} and ~\ref{prop:spectral_series_testing}, in Section~\ref{sec:spectral_projection}, show that under rather general conditions on the design distribution, classical spectral projection methods achieve minimax rates of convergence. Then in Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps}, \ref{sec:out_of_sample} and and~\ref{sec:manifold_adaptivity}, we give various upper bounds on the error of Laplacian eigenmaps (summarized above), as well as summarizing our analysis at a high-level and providing some key intermediary results. In Section~\ref{sec:experiments} we examine the empirical behavior of Laplacian eigenmaps, and show that even at moderate sample sizes Laplacian eigenmaps is competitive with classical spectral projection. We conclude with some discussion in Section~\ref{sec:discussion}, comparing Laplacian eigenmaps to a few selected estimators and commenting on some computational considerations.

\paragraph{Notation.}
We frequently refer to various classical function classes. For an open set $\mc{X}$ equipped with a volume form $d\mu$, we let $L^2(\mc{X})$ denote the set of functions $f$ for which $\|f\|_{L^2(\mc{X})}^2 := \int f^2 \,d\mu  < \infty$, and equip $L^2(\mc{X})$ with the norm $\|\cdot\|_{L^2(\mc{X})}$. We define $\dotp{f}{g}_P := \int fg\,dP$, and let $L^2(P)$ contain those functions $f$ for which $\|f\|_P^2 := \dotp{f}{f}_P$ is finite. Finally, we let $L^2(P_n)$ consist of those ``functions'' $f: \set{X_1,\ldots,X_n} \to \Reals$ for which the empirical norm $\|f\|_{n}^2 := \frac{1}{n}\sum_{i = 1}^{n} \bigl(f(X_i)\bigr)^2 < \infty$. When there is no chance of confusion, we will sometimes associate functions in $L^2(P_n)$ with vectors in $\Reals^n$, and vice versa. We use $C^k(\mc{X})$ to refer to functions which are $k$ times continuously differentiable in $\mc{X}$, either for some integer $k \geq 1$ or for $k = \infty$. We let $C_c^{\infty}(\mc{X})$ represent those functions in $C^{\infty}(\mc{X})$ with support $V$ compactly contained in $\mc{X}$, meaning $\wb{V} \subseteq \mc{X}$. We write $\partial f/\partial r_i$ for the partial derivative of $f$ in the $i$th standard coordinate of $\Rd$, and use the multi-index notation $D^{\alpha}f := \partial^{|\alpha|}f/\partial^{\alpha_1}x_1\ldots\partial^{\alpha_d}x_d$ for multi-indices $\alpha \in \Reals^d$.

We write $\|\cdot\| = \|\cdot\|_2$ for Euclidean norm, $|\cdot| = \|\cdot\|_1$ for $\ell_1$ norm, and $d_{\mc{X}}(x',x)$ for the geodesic distance between points $x$ and $x'$ on a manifold $\mc{X}$. Then for a given $\delta > 0$, $B(x,\delta)$ is the radius-$\delta$ ball with respect to Euclidean distance, whereas $B_{\mc{X}}(x,\delta)$ is the radius-$\delta$ ball with respect to geodesic distance. Letting $T_x(\mc{X})$ be the tangent space at a point $x \in \mc{X}$, we write $B_m(v,\delta) \subset T_x(\mc{X})$ for the radius-$\delta$ ball centered at $v \in T_x(\mc{X})$.

For sequences $(a_n)$ and $(b_n)$, we use the asymptotic notation $a_n \lesssim b_n$ to mean that there exists a number $C$ such that $a_n \leq C b_n$ for all $n$. We write $a_n \asymp b_n$ when $a_n \lesssim b_n$ and $b_n \lesssim a_n$. On the other hand we write $a_n = o(b_n)$ when $\lim a_n/b_n = 0$, and likewise $a_n = \omega(b_n)$ when $\lim a_n/b_n = \infty$. Finally $a \vee b := \max\{a,b\}$ and $a \wedge b := \min\{a,b\}$.

### 2021-08-13 manuscripts/laplacian_eigenmaps/sections/preliminaries.tex
Fixed all of Siva's small changes. Comprehensive changes made to introduction of Laplacian Eigenmaps paper.

\section{Preliminaries}
\label{sec:setup_main_results}

In this section, we precisely define our modeling assumptions and the Laplacian eigenmaps methods we study.

\subsection{Nonparametric regression over Sobolev spaces}
\label{sec:regression_laplacian_eigenmaps}

We will operate in the usual setting of nonparametric regression with random design, in which we observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$. The design points $X_1,\ldots,X_n$ are sampled from a distribution $P$ with support $\mc{X} \subseteq \Rd$, and the responses follow the signal plus noise model
\begin{equation}
\label{eqn:model}
Y_i = f_0(X_i) + w_i,
\end{equation}
with regression function $f_0: \mc{X} \to \Reals$, and $w_i \sim N(0,1)$ independent Gaussian noise. 
% (AG): Removed the comment regarding noise with variance = \sigma^2.

We now formulate two models, which differ in the assumed nature of the support $\mc{X}$ of the design distribution $P$: the \emph{flat Euclidean} and \emph{manifold} models.

\begin{definition}[Flat Euclidean model]
	\label{def:model_flat_euclidean}
	The data $(X_1,Y_1),\ldots,(X_n,Y_n)$ are sampled according to~\eqref{eqn:model}. The support $\mc{X}$ of the design distribution $P$ is an open, connected, and bounded subset of $\Rd$, with Lipschitz boundary. The distribution $P$ admits a Lipschitz density $p$ with respect to the $d$-dimensional Lebesgue measure $\nu$, which is bounded away from $0$ and $\infty$,
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty, \quad \textrm{for all $x \in \mc{X}$.}
	\end{equation*}
\end{definition}

In the following, we recall that the injectivity radius of an $m$-dimensional Riemannian manifold $\mc{X}$ is the maximum value of $\delta$ such that the exponential map $\exp_x: B_m(0,\delta) \subset T_x(\mc{X}) \to B_{\mc{X}}(x,\delta) \subset \mc{X}$ is a diffeomorphism for all $x \in \mc{X}$.
\begin{definition}[Manifold model]
	\label{def:model_manifold}
	The data $(X_1,Y_1),\ldots,(X_n,Y_n)$ are sampled according to~\eqref{eqn:model}. 
	The support $\mc{X}$ of the design distribution $P$ is a closed, connected, smooth and boundaryless Riemannanian manifold embedded in $\Rd$, of intrinsic dimension $1 \leq m < d$. The injectivity radius of $\mc{X}$ is lower bounded by a positive constant $i_0 > 0$. The design distribution $P$ admits a Lipschitz density $p$ with respect to the volume form $d\mu$ induced by the Riemannian structure of $\mc{X}$, which is bounded away from $0$ and $\infty$,
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty, \quad \textrm{for all $x \in \mc{X}$.}
	\end{equation*}
\end{definition}

Finally, at various points we will have to assume that the density $p$ also satisfies different types of higher-order regularity, beyond Lipschitz continuity. When such assumptions are necessary we will state them explicitly.

\subsection{Laplacian eigenmaps}
We now formally define the estimator and test statistic we study. Both are derived from eigenvectors of a graph Laplacian.  For a positive, symmetric kernel $\eta: [0,\infty) \to [0,\infty)$, and a radius parameter $\varepsilon > 0$, let $G = ([n],W)$ be the neighborhood graph formed over the design points $\{X_1,\ldots,X_n\}$, with a weighted edge $W_{ij} = \eta(\|X_i - X_j\|/\varepsilon)$ between vertices $i$ and $j$. Then the 
\emph{neighborhood graph Laplacian} $L_{n,\varepsilon}: \Reals^n \to \Reals$ is defined by its action on vectors $u \in \Reals^n$ as
\begin{equation}
\label{eqn:neighborhood_graph_laplacian}
\bigl(L_{n,\varepsilon}u\bigr)_i := \frac{1}{n\varepsilon^{2 + \mathrm{dim}(\mc{X})}} \sum_{j = 1}^{n} \bigl(u_i - u_j\bigr) \eta\biggl(\frac{\|X_i - X_j\|}{\varepsilon}\biggr).
\end{equation}
(Here $\mathrm{dim}(\mc{X})$ stands for the dimension of $\mc{X}$. It is equal to $d$ under the assumptions of Model~\ref{def:model_flat_euclidean}, and equal to $m$ under the assumptions of Model~\ref{def:model_manifold}. The pre-factor $(n\varepsilon^{2 + \mathrm{dim}(\mc{X})})^{-1}$ is purely for convenience in taking limits as $n \to \infty, \varepsilon \to 0$). Written in standard coordinates we have $(n\varepsilon^{\dim(\mc{X}) + 2}) \cdot L_{n,\varepsilon} = D - W$, where $D \in \Reals^{n \times n}$ is the diagonal degree matrix, $D_{ii} = \sum_{i = 1}^{n} W_{ij}$.

The graph Laplacian is a positive semi-definite matrix, and admits the eigendecomposition $L_{n,\varepsilon} = \sum_{k = 1}^{n} \lambda_k v_k v_k^{\top}$, where for each $k = 1,\ldots,n$ the eigenvalue-eigenvector pair $(\lambda_k,v_k)$ satisfies
\begin{equation*}
L_{n,\varepsilon}v_k = \lambda_k v_k, \quad \|v_k\|_2^2 = 1.
\end{equation*}
We will assume without loss of generality that each eigenvalue $\lambda$ of $L_{n,\varepsilon}$ has algebraic multiplicity $1$, and so we can index the eigenpairs $(\lambda_1,v_1),\ldots,(\lambda_n,v_n)$ in ascending order of eigenvalue, $0 = \lambda_1 < \ldots < \lambda_n$.

The Laplacian eigenmaps estimator $\wh{f}$ simply projects the response vector ${\bf Y}$ onto the first $K$ eigenvectors of $L_{n,\varepsilon}$: letting $V_K \in \Reals^{n \times K}$ be the matrix with columns $v_1,\ldots,v_K$, we have that
\begin{equation}
\label{eqn:laplacian_eigenmaps_estimator}
\wh{f} := \sum_{k = 1}^{K} \dotp{{\bf Y}}{v_k}_{n} v_k = \frac{1}{n} V_K V_K^{\top} {\bf Y}.
\end{equation} 
Thus $\wh{f}$ is equivalently a vector in $\Reals^n$, or a function in $L^2(P_n)$. If $\wh{f}$ is a reasonable estimate of $f_0$, then the Laplacian eigenmaps test statistic
\begin{equation}
\label{eqn:laplacian_eigenmaps_test}
\wh{T} := \|\wh{f}\|_n^2
\end{equation}
is in turn a reasonable estimate of $\|f_0\|_{P}^2$, and can be used to distinguish whether or not $f_0 = 0$.

% AG 8/12/21: Commented out due to overlap with footnote in new introduction.
% It may be helpful to comment briefly on the term ``Laplacian eigenmaps'', which we use a bit differently than is typical in the literature. Laplacian eigenmaps typically refers to an algorithm for embedding, which maps each design point $X_1,\ldots,X_n$ to $\Reals^K$ according to $X_i \mapsto (v_{1,i}, \ldots, v_{K,i})$. Viewing this embedding as a feature map, we can then interpret the estimator $\wh{f}$ as the least-squares solution to a linear regression problem with responses $Y_1,\ldots,Y_n$ and features $v_1,\ldots,v_K$. Often, the Laplacian eigenmaps embedding is viewed as a tool for dimensionality reduction, wherein it is implicitly assumed that $K$ is much smaller than $d$. We will neither explicitly nor implicitly take $K < d$; indeed, the embedding perspective is not particularly illuminating in what follows, and we do not henceforth make reference to it. Instead, we use ``Laplacian eigenmaps'' to directly refer to the estimator $\wh{f}$ or test statistic $\wh{T}$.

\subsection{Sobolev Classes}
\label{sec:sobolev}
We now review the definition of Sobolev classes, dividing our discussion into two cases, the flat Euclidean case (Model~\ref{def:model_flat_euclidean}) and the manifold case (Model~\ref{def:model_manifold}).

\paragraph{Flat Euclidean case.}
Take $\mc{X}$ to be a open, connected and bounded set with Lipschitz boundary, as in Model~\ref{def:model_flat_euclidean}. Recall that for a given multi-index $\alpha \in \mathbb{N}^d$, a function $f$ is \emph{$\alpha$-weakly differentiable} if there exists some $h \in L^1(\mc{X})$ such that
\begin{equation*}
\int_{\mc{X}} h g = (-1)^{|\alpha|} \int_{\mc{X}} f D^{\alpha}g, \quad \textrm{for every $g \in C_c^{\infty}(\mc{X})$.}
\end{equation*}
If such a function $h$ exists, it is the $\alpha$th weak partial derivative of $f$, and denoted by $D^{\alpha}f := h$. For functions $f$ which are $|\alpha|$-times differentiable, this coincides with the classical definition of derivative.

\begin{definition}[Sobolev space on a flat Euclidean domain]
	\label{def:sobolev_space}
	For an integer $s \geq 1$, a function $f \in L^2(\mc{X})$ belongs to the Sobolev space $H^s(\mc{X})$ if for all $|\alpha| \leq s$, the weak derivatives $D^{\alpha}f$ exists and satisfy $D^{\alpha}f \in L^2(\mc{X})$. The $j$th order semi-norm for $f \in H^s(\mc{X})$ is $|f|_{H^j(\mc{X})} := \sum_{|\alpha| = j}\|D^{\alpha}f\|_{\Leb^2(\mc{X})}$, and the corresponding norm
	\begin{equation*}
	\|f\|_{H^s(\mc{X})}^2 := \|f\|_{\Leb^2(\mc{X})}^2 + \sum_{j = 1}^{s} |f|_{H^j(\mc{X})}^2,
	\end{equation*}
	induces the Sobolev ball
	\begin{equation*}
	H^s(\mc{X};M) := \bigl\{f \in H^s(\mc{X}): \|f\|_{H^s(\mc{X})} \leq M\bigr\}.
	\end{equation*} 
\end{definition}
We note that $H^s(\mc{X})$ is the completion of $C^{\infty}(\mc{X})$ with respect to the $\|\cdot\|_{H^s(\mc{X})}$ norm, so that $C^{\infty}(\mc{X})$ is dense in $H^s(\mc{X})$.

\paragraph{Manifold case.}

There are several equivalent ways to define Sobolev spaces on a compact, smooth, $m$-dimensional Riemannian manifold embedded in $\Reals^d$. We will stick with a definition that parallels our setup in the flat Euclidean setting as much as possible. To do so, we first recall the notion of partial derivatives on a manifold, which are defined with respect to a local coordinate system. Letting $r_1,\ldots,r_m$ be the standard basis of $\Reals^m$, for a given chart $(\phi,U)$ (meaning an open set $U \subseteq \mc{X}$, and a smooth mapping $\phi: U \to \Reals^m$) we write $\phi =: (x_1,\ldots,x_m)$ in local coordinates, meaning $x_i = r_i \circ \phi$. Then the partial derivative $\partial f/\partial x_i$ of a function $f$ with respect to $x_i$ at $x \in U$ is
\begin{equation*}
\frac{\partial f}{\partial x_i}(x) := \frac{\partial(f \circ \phi^{-1})}{\partial r_i}\bigl(\phi(x)\bigr).
\end{equation*}
The right hand side should be interpreted in the weak sense of derivative. As before, we use the multi-index notation $D^{\alpha}f := \partial^{|\alpha|}f/\partial^{\alpha_1}x_1\ldots\partial^{\alpha_m}x_m$.

\begin{definition}[Sobolev space on a manifold]
	\label{def:sobolev_space_manifold}
	A function $f \in \Leb^2(\mc{X})$ belongs to the Sobolev space $H^{s}(\mc{X})$ if for all $\abs{\alpha} \leq s$, the weak derivatives $D^{\alpha}f$ exist and satisfy  $D^{\alpha}f \in \Leb^2(\mc{X})$. The $j$th order semi-norm $|f|_{H^j(\mc{X})}$, the norm $\|f\|_{H^s(\mc{X})}$, and the ball $H^s(\mc{X};M)$ are all defined as in Definition~\ref{def:sobolev_space}.
\end{definition}
The partial derivatives $D^{\alpha}f$ will depend on the choice of local coordinates, and so will the resulting Sobolev norm~$\|f\|_{H^s(\mc{X})}$. However, the important point is that regardless of the choice of local coordinates the resulting norms will be equivalent\footnote{Recall that norms $\|\cdot\|_1$ and $\|\cdot\|_2$ on a space $\mc{F}$ are said to be equivalent if there exist constants $c$ and $C$ such that
\begin{equation*}
c \|f\|_1 \leq \|f\|_2 \leq C \|f\|_1 \quad \textrm{for all $f \in \mc{F}$.}
\end{equation*}} 
and so the ultimate Sobolev space $H^s(\mc{X})$ is independent of the choice of local coordinates.

\paragraph{Boundary conditions.}
In the flat Euclidean model (Model~\ref{def:model_flat_euclidean}), in order to show that Laplacian eigenmaps is optimal over $H^{s}(\mc{X})$ for $s > 1$ we will need to assume that the regression function $f_0$ satisfies some boundary conditions. In particular, we will assume that $f_0$ is zero-trace.
\begin{definition}[Zero-trace Sobolev space]
	\label{def:zero_trace_sobolev_space}
	The \emph{order-s zero-trace Sobolev space} $H_0^s(\mc{X})$ is the closure of $C_c^\infty(\mc{X})$ with respect to $\|\cdot\|_{H^s(\mc{X})}$ norm. That is, $f \in H_0^s(\mc{X})$ if $f \in H^s(\mc{X})$ and additionally there exists a sequence $f_1,f_2,\ldots$ of functions in $C_c^{\infty}(\mc{X})$ such that
	\begin{equation*}
	\lim_{k \to \infty}\|f_k - f\|_{H^s(\mc{X})} = 0.
	\end{equation*}
	The normed ball $H_0^{s}(\mc{X};M) := H_0^{s}(\mc{X}) \cap H^{s}(\mc{X};M)$.
\end{definition}
The zero-trace condition can be made more concrete when $f \in C^\infty(\mc{X})$, since we can then speak of the pointwise behavior of $f$ and its derivatives. Letting $\partial/(\partial {\bf n})$ be the partial derivative operator in the direction of the vector $\mathbf{n}$ normal to the boundary of $\mc{X}$, then the zero-trace condition implies that $\partial^{k}f/\partial{\bf n}^k(x) = 0$ for each $k = 0,\ldots,s - 1$, and for all $x \in \partial\mc{X}$.

% (SB via AG) Move following chunk to appendix, and replace with one sentence explaining that when studying spectral projection methods, it is standard to impose boundary conditions.
We now explain why Laplacian eigenmaps should be optimal only when $f_0$ satisfies certain boundary conditions. Let $(\lambda_1(\Delta_P),\psi_1),(\lambda_2(\Delta_P),\psi_2),\ldots$ be the solutions to the weighted Laplacian eigenvector equation with Neumann boundary conditions, 
\begin{equation}
\label{eqn:laplacian_eigenfunction_neumann}
\Delta_P\psi_k = \lambda_{k}(\Delta_P) \psi_k, \quad \frac{\partial}{\partial {\bf n}}\psi_k = 0~~\textrm{on $\partial \mc{X}$}.
\end{equation}
As we have already alluded to, it is known~\citep{garciatrillos18} that each graph Laplacian eigenpair ($\lambda_k$,$v_k$) converges to a corresponding solution ($\lambda_k(\Delta_P)$, $\psi_k$) of~\eqref{eqn:laplacian_eigenfunction_neumann}. Thus it is relevant to consider which Sobolev functions $f \in H^s(M)$ we can reconstruct using the eigenfunctions $\psi_1,\psi_2,\ldots$. To that end, we introduce the spectrally defined Sobolev space
\begin{equation}
\label{eqn:spectral_sobolev_space}
\mc{H}^s(\mc{X}) := \biggl\{f \in L^2(\mc{X}): \sum_{k = 1}^{\infty} \bigl[\dotp{f}{\psi_k}_P\bigr]^2 \cdot\bigl[\lambda_k(\Delta_P)\bigr]^s < \infty \biggr\}.
\end{equation}
Under the conditions $p \in C^{\infty}(\mc{X})$ and $\partial\mc{X} \in C^{1,1}$, \cite{dunlop2020} show the strict inclusion $\mc{H}^{2s}(\mc{X}) \subset H^{2s}(\mc{X})$. More precisely, they show that for any $s > 0$,
\begin{equation}
\label{eqn:spectral_sobolev_space_equivalence}
\mc{H}^{2s}(\mc{X}) = \biggl\{f \in H^{2s}(\mc{X}): \frac{\partial \Delta_P^rf}{\partial {\bf n}} = 0~\textrm{on}~\partial\mc{X},~~\textrm{for all $0 \leq r \leq s - 1$} \biggr\},
\end{equation}
and for any $s \geq 0$, $\mc{H}^{2s + 1}(\mc{X}) = \mc{H}^{2s}(\mc{X}) \cap H^{2s + 1}(\mc{X})$. If $P$ is uniform on $\mc{X}$, then~\eqref{eqn:spectral_sobolev_space_equivalence} means that a Sobolev function $f \in H^{s}(\mc{X})$ additionally belongs to $\mc{H}^{s}(\mc{X})$ only if all its odd lower order derivatives vanish at the boundary $\partial\mc{X}$.

The bottom line is that the eigenvectors $v_k$ of the graph Laplacian accurately approximate only those functions $f \in H^{s}(\mc{X})$ which satisfy some additional boundary conditions. Although the zero-trace boundary condition is more restrictive than~\eqref{eqn:spectral_sobolev_space_equivalence}---since it also requires that derivatives of even-order be equal to $0$---the point is that some kind of boundary condition will be necessary in order to obtain optimal rates. We will stick to the zero-trace condition, since it greatly eases some of the steps in our proofs.
% (SB via AG) Move above chun to appendix.

On the other hand, in the manifold model (Model~\ref{def:model_manifold}) the domain $\mc{X}$ is without boundary---precisely, every point $x \in \mc{X}$ has a neighborhood that is homeomorphic to an open set in $\Reals^m$, for instance the open ball $B(x,\delta)$ for any $\delta$ smaller than the injectivity radius $i_0$---and so boundary conditions are irrelevant.

\subsection{Minimax Rates and Spectral Projection Methods}
\label{subsec:minimax_rates_sobolev}
We now review the minimax estimation and goodness-of-fit testing rates over Sobolev balls. We will pay special attention to certain classical spectral projection methods which achieve these rates. This is because, as we have already discussed, classical spectral projection methods are very related to Laplacian eigenmaps.

\paragraph{Estimation rates over Sobolev balls.}
In the estimation problem, we ask for an estimator---formally, a Borel measurable function $\wh{f}$ that maps from $\mc{X}$ to $\Reals$---which is close to the regression function $f_0$ with respect to the squared norm $\|\wh{f} - f_0\|_{P}^2$. Under Model~\ref{def:model_flat_euclidean}, the minimax estimation rate over the order-$s$ Sobolev ball is
\begin{equation}
\label{eqn:sobolev_space_minimax_estimation_rate}
\inf_{\wh{f}} \sup_{f_0} \Ebb\|\wh{f} - f_0\|_P^2 \asymp M^2(M^2n)^{-2s/(2s + d)};
\end{equation}
here the infimum is taken over all estimators $\wh{f}$, and the supremum over all $f_0 \in H^1(\mc{X};M)$ (first-order case) or $f_0 \in H_0^s(\mc{X};M)$ (higher-order case), and we assume $n^{-1/2} \lesssim M$.
% (SB via AG) Supremum also taken over all densities p? Do the densities have to satisfy certain conditions?

% (SB via AG) 
% 	1. Not clear that $\wt{f}$ is an estimator in Model 2.1.
%   2. My description of the analysis sells the proposition short.
%   3. Replace the proof description with the proof itself.
The lower bound in~\eqref{eqn:sobolev_space_minimax_estimation_rate} is due to~\citep{stone1980} (at least for the case of $M$ constant, as is most typically considered). The upper bound can be certified by a particular spectral projection estimator $\wt{f}$,
\begin{equation}
\label{eqn:spectral_series_estimator}
\wt{f} := \sum_{k = 1}^{K} \dotp{Y}{\psi_k}_n \psi_k,
\end{equation}
where $\psi_k$ are the eigenfunctions of $\Delta_P$ defined in~\eqref{eqn:laplacian_eigenfunction_neumann}. The optimality of spectral projection estimators over Sobolev type spaces is generally well-understood. For instance, see \cite{tsybakov08,johnstone2011,gine16}, who work in the Gaussian sequence model and show that analogous estimators are optimal over Sobolev ellipsoids. However, we have not found an analysis of the specific estimator~\eqref{eqn:spectral_series_estimator} under Model~\ref{def:model_flat_euclidean}, and so for completeness we state the result in the following proposition.
\begin{proposition}
	\label{prop:spectral_series_estimation}
	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $\partial \mc{X} \in C^{1,1}$, that $p \in C^{\infty}(\mc{X})$, and that $f_0 \in H^1(\mc{X};M)$ (first-order case) or $f_0 \in H_0^s(\mc{X};M)$ for some $s > 1$ (higher-order case). Then there exists a constant $C$ which does not depend on $f_0,M$ or $n$ such that the following statement holds: if the spectral projection estimator $\wt{f}$ is computed with parameter $K = \floor{M^2n}^{d/(2s + d)}$, then
	\begin{equation*}
	\Ebb\bigl[\|\wt{f} - f_0\|_P^2\bigr] \leq C \min\bigl\{M^2(M^2n)^{-2s/(2s + d)}, M^2\bigr\}
	\end{equation*}
\end{proposition}
The proof of Proposition~\ref{prop:spectral_series_estimation}, along with proofs of all of our results, can be found in the appendix. It is worth mentioning some aspects of the analysis here, because it sets the stage for the strategy we will use to analyze Laplacian eigenmaps.

In particular, three essential facts are needed to establish Proposition~\ref{prop:spectral_series_estimation}. 
\begin{enumerate}
	\item The continuous embedding of $H_0^s(\mc{X})$ into $\mc{H}^s(\mc{X})$---recall the latter is defined in~\eqref{eqn:spectral_sobolev_space}---which is a consequence of the zero-trace condition, the conditions on $\partial \mc{X}$ and $p$, and~\eqref{eqn:spectral_sobolev_space_equivalence}.
	\item Weyl's Law, which gives the asymptotic scaling of eigenvalues $\lambda_{k}(\Delta_P) \asymp k^{2/d}$, and allows us to properly control the bias induced by spectral projection.
	\item A local version of Weyl's law, which gives an estimate on $\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2$, and allows us to appropriately control the difference $\dotp{f_0}{\psi_k}_n - \dotp{f_0}{\psi_k}_P$  between the empirical and population Fourier coefficients.
\end{enumerate}
With these facts in hand the proof of Proposition~\ref{prop:spectral_series_estimation} follows from calculations standard to analysis of the Gaussian sequence model. Our analysis of Laplacian eigenmaps will depend on analogues of the first and second of these facts, with the space $\mc{H}^s(\mc{X})$ and eigenvalues $\lambda_k(\Delta_P)$ replaced by alternatives suitably defined with respect to the neighborhood graph Laplacian $L_{n,\varepsilon}$.

\paragraph{Goodness-of-fit testing rates over Sobolev balls.}
In the goodness-of-fit testing problem, we ask for a test function---formally, a Borel measurable function $\phi$ that takes values in $\{0,1\}$--- which can distinguish between the hypotheses
\begin{equation}
\mathbf{H}_0: f_0 = f_0^{\ast}, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{f_0^{\ast}\}.
\end{equation} 
Typically, the null hypothesis $f_0 = f_0^{\ast} \in \mc{F}$ reflects the absence of interesting structure, and $\mc{F} \setminus  \{f_0^{\ast}\}$ is a set of smooth departures from this null. To fix ideas, as in \citet{ingster2009} we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\ast} = 0$ and $\mc{F} = \mc{H}^s(\mc{X};M)$ is a Sobolev ball. This is without loss of generality since our test statistic and its analysis are easily modified to handle the case when $f_0^{\ast}$ is not $0$, by simply subtracting $f_0^{\ast}(X_i)$ from each observation $Y_i$.

The Type I error of a test $\phi$ is $\mathbb{E}_0[\phi]$, and if $\mathbb{E}_0[\phi] \leq a$ for a given $a \in (0,1)$ we refer to $\phi$ as a level-$a$ test\footnote{We reserve the more common symbol $\alpha$ for multi-indices, so as to avoid confusion}. The worst-case risk of $\phi$ over $\mc{F}$ is
\begin{equation*}
R_n(\phi, \mc{F}, \epsilon) := \sup\Bigl\{\mathbb{E}_{f_0}[1 - \phi]: f_0 \in \mc{F}, \|f_0\|_{P} > \epsilon\Bigr\},
\end{equation*}
and for a given constant $b \in (0,1)$, the minimax critical radius $\epsilon_n(\mc{F})$ is the smallest value of $\epsilon$ such that some level-$a$ test has worst-case risk of at most $b$. Formally,
\begin{equation*}
\epsilon_n(\mc{F}) := \inf\Bigl\{\epsilon > 0: \inf_{\phi} R_n(\phi,\mc{F},\epsilon) \leq b \Bigr\},
\end{equation*} 
where in the above the infimum is over all level-$a$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Clearly, the minimax critical radius $\epsilon_n(\mc{F})$ depends on $a$ and $b$. However, we adopt the typical convention of treating $\alpha,b \in (0,1)$ and  as small but fixed positive constants; hence they will not affect the testing error rates, and we suppress them notationally.} We will refer to the rate at which the squared critical radius $\epsilon_n^2(\mc{F})$ tends to $0$ as the minimax testing rate.

Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing rate is much smaller than the minimax estimation rate. \citet{ingster2009} give the minimax testing rate in the special case where $\mc{X} = [0,1]^d$ and $P$ is uniform, and we restate their result in the terms and notation of our paper.
\begin{theorem}[Theorem 1 of \citet{ingster2009}]
	\label{thm:spectral_series_testing}
	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $\mc{X} = [0,1]^d$ and $P$ is the uniform distribution on $[0,1]^d$. Then there exist constants $c,C,n_0$ which do not depend on $f_0$ or $n$ such that for all $n \geq n_0$,
	\begin{equation}
	\label{eqn:sobolev_space_testing_critical_radius}
	c n^{-4s/(4s + d)} \leq \epsilon_n^2\bigl(\mc{H}^s(\mc{X};1)\bigr) \leq C n^{-4s/(4s + d)}~~\textrm{for $1 \leq d < 4s$.}
	\end{equation}
\end{theorem}
The analysis used by \citet{ingster2009} to show the upper bound in~\eqref{eqn:sobolev_space_testing_critical_radius} relies on a similar trio of facts as used in the proof of Proposition~\ref{prop:spectral_series_estimation}. It is otherwise reminiscent of calculations made in the Gaussian sequence model, which can be found in~\cite{ingster2012}. This analysis can be straightforwardly adapted to handle design distributions $P$ that satisfy the conditions of Model~\ref{def:model_flat_euclidean}, or to handle the case where $M$ is not $1$. Finally, the test \citet{ingster2009} use to certify the upper bound in~\eqref{eqn:sobolev_space_testing_critical_radius} is implicitly a spectral projection method: the test statistic is the $L^2(P_n)$ norm of a particular spectral projection estimator.\footnote{\cite{ingster2009} project the responses onto a the span of trigonometric basis functions. Since $P$ is uniform on the unit cube, such functions are eigenfunctions of $\Delta_P$ when the right boundary conditions are imposed.}

A major difference between testing and estimation over Sobolev spaces is the requirement that $4s > d$. When $4s \leq d$, the functions in $H^s(\Xset)$ are very irregular. Crucially, $H^s(\Xset)$ no longer continuously embeds into $\Leb^4(\Xset)$ when $4s \leq d$, and test statistics using the $L^2(P_n)$ norm are no longer guaranteed to have finite variance.\footnote{Note that this will not affect the analysis for estimation, because for estimation we only need to control the first two moments of $f_0$.} In fact, we have not seen any analysis which describes the minimax rate for nonparametric regression testing over Sobolev spaces in the $4s \leq d$ regime. However, if one explicitly assumes that $f_0 \in \Leb^4(\mc{X})$, then the critical radius is characterized by the dimension-free rate $\epsilon_n^2(\Leb^4(\mc{X};M)) \asymp Mn^{-1/2}$.\footnote{As a sanity check, note that when $1 \leq d < 4s$, $n^{-1/2}$ converges to $0$ at a slower rate than $n^{-4s/(4s + d)}$, which is the minimax rate for $M = 1$ given in~\eqref{eqn:sobolev_space_testing_critical_radius}.} As we discuss after our first main theorem regarding testing with Laplacian eigenmaps (Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}), this rate is achievable by a test based on $\wh{T}$.

\paragraph{Manifold setup.}
Under Model~\ref{def:model_manifold}, both~\eqref{eqn:sobolev_space_minimax_estimation_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced everywhere by the intrinsic dimension $m$. The estimator $\wt{f}$ and a test using the statistic $\wt{T}$---with $\psi_1,\psi_2,\ldots$ now the eigenfunctions of the manifold weighted Laplace-Beltrami operator $\Delta_P$---achieve the optimal estimation and testing rates. This is because each of the three facts mentioned after Proposition~\ref{prop:spectral_series_estimation} have analogues when the domain is a smooth manifold (See~\citet{hendriks1990}, who analyzes a spectral projection density estimator, for details.)

\paragraph{In-sample mean squared error.}
As mentioned in our introduction, roughly speaking one of our main conclusions is that the Laplacian eigenmaps estimator $\wh{f}$ is minimax rate-optimal. It is worth being clear about what we do and do not mean by this statement. We do not mean that the estimator $\wh{f}$ will match the upper bound given in~\eqref{eqn:sobolev_space_minimax_estimation_rate}, since such a statement does not make sense when the estimator is defined only at the random design points $X_1,\ldots,X_n$. Instead we will measure loss using the squared $L^2(P_n)$ error. In Section~\ref{sec:out_of_sample} we show that an extension of $\wh{f}$ defined over all $\mc{X}$ has $L^2(P)$ error comparable to the $L^2(P_n)$ error of $\wh{f}$. We also believe that in the random design setting we work in, simple arguments will imply that $L^2(P_n)$ risk has the same minimax rate of convergence as $L^2(P)$ risk. We sketch such an argument in Section~\ref{sec:out_of_sample}, but do not further pursue the details.

% (SB via AG): Move this somewhere later.
Additionally, we will not actually measure accuracy using the expectation of the loss. Rather, we will give a constant probability bound on $\|\cdot\|_n^2$. For instance, when $f_0 \in H^1(\mc{X};1)$, we will show that with probability $1 - \delta$ the loss $\|\wh{f} - f_0\|_n^2 \leq C_{\delta} n^{-2/(2 + d)}$, for a constant $C_{\delta}$ that depends on $\delta$ but not on $f_0$ or $n$. Thus we give an upper bound on the $(1 - \delta)$th quantile of $\|\wh{f} - f_0\|_n^2$, rather than an upper bound on its expectation. We explain the reason for this in Section~\ref{sec:minimax_optimal_laplacian_eigenmaps}. We also show that if $f_0$ is bounded in a larger norm---for instance, if it is H\"{o}lder rather than Sobolev smooth---then we can obtain bounds on the expected $L^2(P_n)$ loss.

There is one other subtlety introduced by the use of in-sample mean squared error. Technically speaking, elements $f \in H^s(\mc{X})$ are equivalence classes, defined only up to a set of measure zero. Thus one cannot speak of the pointwise evaluation $f_0(X_i)$, as we do by defining our target of estimation to be $f_0(X_i)$, $i=1,\ldots,n$, until one selects \emph{representatives}. When $s > d/2$, every element $f$ of $H^s(\mc{X})$ admits a continuous version $f^{\ast}$, and as is standard we set this to be our favored representative. When $s \leq d/2$, some elements in $H^s(\mc{X})$ do not have any continuous version; however they admit a \emph{quasi-continuous} version \citep{evans15} known as the \emph{precise representative}, and we use this representative. To be clear, however, it does not really matter which representative we choose. Since all versions agree except on a set of measure zero, and since $P$ is absolutely continuous with respect to Lebesgue measure (in Model~\ref{def:model_flat_euclidean}) or the volume form $d\mu$ (in Model~\ref{def:model_manifold}), with probability $1$ any two versions $g_0, h_0 \in f_0$ will satisfy $g_0(X_i) = h_0(X_i)$ for all $i = 1,\ldots,n$. The bottom line is that we can use the notation $f_0(X_i)$ without fear of ambiguity or confusion.

Finally, we note that for testing none of these comments are relevant. We will show that our test has small worst-case risk whenever $\epsilon \gtrsim \epsilon_n(H_0^s(\mc{X};M))$, thus establishing that it is a minimax optimal test in the usual sense.

### 2021-08-23 manuscripts/laplacian_eigenmaps/appendix.tex
Modest change to appendix.
Updated appendix.
Updated preliminaries section of Laplacian Eigenmaps manuscript.

@@
 \end{equation*}
-
-\section{Analysis of Spectral Series Estimator}
-\label{sec:analysis_spectral_series_estimator}
-In this section we prove Proposition~\ref{prop:spectral_series_estimation}.
-As mentioned in our main text, this proof relies on three facts: (i) a continuous embedding of $H_0^s(\mc{X})$ into $\mc{H}^s(\mc{X})$, (ii) the Weyl's Law asymptotic scaling $\lambda_k(\Delta_P) \asymp k^{2/d}$, and (iii) a local Weyl's Law scaling $\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2 \lesssim K$.
-(Recall that $(\lambda_k(\Delta_P),\psi_k)$ are defined to be solutions to~\eqref{eqn:laplacian_eigenfunction_neumann}).
-We record the precise results we need in the following three lemmas.
-
-The first result follows immediately from Lemma~17 of \textcolor{blue}{(Dunlop)}.
-Recall the spectral Sobolev class $\mc{H}^s(\mc{X})$ defined in~\eqref{eqn:spectral_sobolev_space}.
-Equip $\mc{H}^s(\mc{X})$ with the norm
-\begin{equation*}
-\|f\|_{\mc{H}^s(\mc{X})}^2 := \sum_{k = 1}^{\infty} \bigl[\dotp{f}{\psi_k}_P\bigr]^2 \cdot [\lambda_k(\Delta_P)]^2.
-\end{equation*}
-\begin{lemma}[cf. Lemma~17 of \textcolor{blue}{(Dunlop)}]
-	\label{lem:zero_trace_spectral_sobolev}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $p \in C^{\infty}(\mc{X})$ and $\partial \mc{X} \in C^{1,1}$.
-Then for any $s \in \mathbb{N}$, we have $H_0^s(\mc{X}) \subseteq \mc{H}^s(\mc{X})$.
-Moreover, there exists a constant $C$ such that for any $f \in H_0^s(\mc{X})$ we have
-	\begin{equation}
-	\label{eqn:zero_trace_spectral_sobolev}
-	\|f\|_{\mc{H}^s(\mc{X})} \leq C\|f\|_{H^s(\mc{X})}.
-	\end{equation}
-\end{lemma}
-
-The particular version of Weyl's Law we use is also due to \textcolor{blue}{(Dunlop)}. 
-\begin{lemma}[cf. Lemma 28 of \textcolor{blue}{(Dunlop)}]
-	\label{lem:weyl}
-	Suppose Model~\ref{def:model_flat_euclidean}. Then there exist positive constants $c$ and $C$ such that for all $k \geq 2$, 
-	\begin{equation}
-	\label{eqn:weyl}
-	ck^{2/d} \leq \lambda_k(\Delta_P) \leq Ck^{2/d}.
-	\end{equation}
-\end{lemma}
-As we discuss in our main text, the asymptotic scaling $\lambda_k(\Delta_P) \asymp k^{2/d}$ also plays a key role in our analysis of Laplacian eigenmaps.
-We point out that although Lemma 28 of \textcolor{blue}{(Dunlop)} assumes $p \in C^{\infty}(\mc{X})$ and $\partial \mc{X} \in C^{1,1}$, the proof uses only the hypotheses of Model~\ref{def:model_flat_euclidean}: that $p$ is bounded away from $0$ and $\infty$, and that $\partial \mc{X}$ is Lipschitz.
-
-Finally, we need a local version of Weyl's Law, due to \textcolor{blue}{(H\"{o}rmander)}.
-\begin{lemma}[cf. Theorem 17.5.3 of \textcolor{blue}{H\"{o}rmander}]
-	\label{lem:local_weyl}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $p \in C^{\infty}(\mc{X})$ and $\partial \mc{X} \in C^{1,1}$.
-Then there exists a positive constant $C$ such that for all $K \in \mathbb{N} \setminus {0}$,
-	\begin{equation}
-	\label{eqn:local_weyl}
-	\sup_{x \in \mc{X}}\biggl\{\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2\biggr\} \leq CK.
-	\end{equation}
-\end{lemma}
-Translated into our notation, Theorem 17.5.3 of \textcolor{blue}{(H\"{o}rmander)} says that
-\begin{equation*}
-\sup_{x \in \mc{X}}\biggl\{\sum_{k: \lambda_k(\Delta_P) \leq \lambda_K(\Delta_P)}  \bigl(\psi_k(x)\bigr)^2\biggr\} \leq C \bigl[\lambda_K(\Delta_P)\bigr]^{d/2},
-\end{equation*}
-and from here~\eqref{eqn:local_weyl} follows from~\eqref{eqn:weyl}.
-\textcolor{blue}{(H\"{o}rmander)} proves the result only when $\partial \mc{X} \in C^{\infty}$, but the proof goes through unchanged when $\partial \mc{X} \in C^1$, since $\partial \mc{X} \in C^1$ is sufficient to apply the relevant Sobolev embedding theorem (Lemma 17.5.2 of~\textcolor{blue}{(H\"{o}rmander)}) that is the key to proving Lemma~\ref{lem:local_weyl}.
-
-We note that the assumptions $p \in C^{\infty}(\mc{X})$ and $\partial\mc{X} \in C^{1,1}$ used in Lemmas~\ref{lem:zero_trace_spectral_sobolev} and~\ref{lem:local_weyl} can likely be removed.
-This would allow us to remove the same assumptions in the statement of Proposition~\ref{prop:spectral_series_estimation}.
-Since this does not change the main point of the proposition, we do not pursue the details further.
-
-\paragraph{Proof (of Proposition~\ref{prop:spectral_series_estimation}).}
-We decompose risk into squared bias and variance,
-\begin{equation}
-\label{pf:spectral_series_estimation_0}
-\Ebb \|\wt{f} - f_0\|_P^2 = \Ebb\| \Ebb[\wt{f}]  - f_0\|_P^2 + \Ebb\| \wt{f} - \Ebb[\wt{f}]\|_P^2.
-\end{equation}
-Then some standard arguments (which we give below) give the following upper bound:
-\begin{equation}
-\label{pf:spectral_series_estimation_1}
-\Ebb \|\wt{f} - f_0\|_P^2 \leq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\bigl[\lambda_{K + 1}(\Delta_P)\bigr]^s} + \frac{K}{n} + \frac{1}{n}\Ebb\Bigl[(f_0(X))^2 \cdot \sum_{k = 1}^{K} (\psi_k(X))^2\Bigr].
-\end{equation}
-The claim of the proposition follows from~\eqref{pf:spectral_series_estimation_1} and Lemmas~\ref{lem:zero_trace_spectral_sobolev}-\ref{lem:local_weyl}, upon taking $K = \floor{(M^2n)^{d/(2s + d)}}$.
-(Here we use the convention $\sum_{k = 1}^{0}c_k = 0$ for any sequence $(c_k)$.) \qed
-
-The upper bound~\eqref{pf:spectral_series_estimation_1} is to be compared with~\eqref{eqn:estimation_biasvariance}, which gives upper bounds on the (design-dependent) squared bias and variance of Laplacian eigenmaps.
-We see~\eqref{eqn:estimation_biasvariance} has terms analogous to the first two terms on the right hand side of~\eqref{pf:spectral_series_estimation_1}, but not the third.
-The relevance of this third term to $L^2$ estimation problems is discussed by~\cite{birge2008}.
-
-\paragraph{Proof (of \eqref{pf:spectral_series_estimation_1}).}
-Since $\{\psi_k\}$ are an orthonormal basis of $L^2(P)$, and $f_0 \in H_0^{s}(\mc{X}) \subseteq L^2(P)$, we have $f_0 = \sum_{k = 1}^{\infty} \dotp{f_0}{\psi_k}_P \psi_k$ in $L^2(P)$.
-This allows us to expand the squared bias in terms of squared Fourier coefficients of $f_0$, and gives the following upper bound,
-\begin{equation*}
-\|f_0 - \Ebb{\wt{f}}\|_P^2 = \sum_{k = K + 1}^{\infty}  \dotp{f_0}{\psi_k}^2 \leq  \frac{1}{\lambda_{K + 1}(\Delta_P)^s} \sum_{k = K + 1}^{\infty} \lambda_{k + 1}(\Delta_P)^s \dotp{f_0}{\psi_k}^2 \leq \frac{\|f_0\|_{\mc{H}^s}}{\bigl[\lambda_{K + 1}(\Delta_P)\bigr]^s}.
-\end{equation*}
-On the other hand, the variance term can be written as the sum of the variance of each empirical Fourier coefficient,
-\begin{equation}
-\label{pf:spectral_series_estimation_2}
-\Ebb\| \wt{f} - \Ebb[\wt{f}]\|_P^2 = \sum_{k = 1}^{K} \Var\Bigl[\dotp{{\bf Y}}{\psi_k}_n\Bigr] 
-\end{equation}
-Then from the law of total variance,
-\begin{equation*}
- \Var\bigl[\dotp{{\bf Y}}{\psi_k}_n\bigr] = \Var\Bigl[\Ebb[\dotp{Y}{\psi_k}_n|{\bf X}\Bigr] + \Ebb\Bigl[\Var[\dotp{Y}{\psi_k}_n|{\bf X}\Bigr] = \Var\Bigl[\dotp{f_0}{\psi_k}_n\Bigr] + \frac{1}{n}\Ebb\Bigl[\|\psi_k\|_n^2\Bigr].
-\end{equation*}
-Finally, $\Ebb \|\psi_k\|_n^2 = \|\psi_k\|_P^2 = 1$, 
-\begin{equation*}
-\Var\Bigl[\dotp{f_0}{\psi_k}_n\Bigr] = \frac{1}{n} \Var\Bigl[f_0(X)\psi_k(X)\Bigr] \leq \frac{1}{n} \Ebb\Bigl[\Bigl(f_0(X)\psi_k(X)\Bigr)^2\Bigr],
-\end{equation*}
-and plugging back into~\eqref{pf:spectral_series_estimation_2} yields the claim.
 

### 2021-08-23 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Updated preliminaries section of Laplacian Eigenmaps manuscript.

@@
 \end{theorem}
-From~\eqref{eqn:laplacian_eigenmaps_estimation_fo} it follows immediately that with constant probability $\|\wh{f} - f_0\|_n^2 \lesssim M^2(M^2n)^{-2/(2 + d)}$ whenever $n^{-1/2} \lesssim M \lesssim n^{1/d}$.
+From~\eqref{eqn:laplacian_eigenmaps_estimation_fo} it follows immediately that with constant probability $\|\wh{f} - f_0\|_n^2 \lesssim M^2(M^2n)^{-2/(2 + d)}$ whenever $n^{-1/2} \lesssim M \lesssim n^{1/d}$, matching the standard minimax estimation rate over Sobolev classes.
 
@@
 Under the stronger assumption that $f_0 \in C^1(\mc{X};M)$ we can replace the factor of $\delta$ by the sharper $\delta^2/n$, which is less than $\delta$ because $\delta \in (0,1)$.
-Then a routine calculation, which we carry out in Appendix~\ref{subsec:upper_bound_l2pn_risk}, shows that the expected $L^2(P_n)$ loss is on the same order as~\eqref{eqn:laplacian_eigenmaps_estimation_fo}, matching the minimax rate.
+Then a routine calculation shows that the expected $L^2(P_n)$ loss is on the same order as~\eqref{eqn:laplacian_eigenmaps_estimation_fo}, matching the minimax rate.
 \end{itemize}
@@
 We proceed to establish suitable upper and lower bounds on these quantities.
+% (AG 8/19/21): Need to include some discussion of how the analysis contrasts with that of the classical spectral projection estimator and test.
 
@@
 	\label{eqn:radius_estimation_manifold}
-	C_0\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/m}, n^{-1/(2(s - 1) + m)}\biggr\} \leq \varepsilon \leq \min\{i_0, K^{-1/m}\},
-	\end{equation}
-	where we recall that $i_0$ is a lower bound on the injectivity radius of $\mc{X}$. Additionally,
+	C_0\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/m}, n^{-1/(2(s - 1) + m)}\biggr\} \leq \varepsilon \leq \min\{\mathrm{inj}(\mc{X}), K^{-1/m}\},
+	\end{equation}
+	where we recall that $\mathrm{inj}(\mc{X})$ is a lower bound on the injectivity radius of $\mc{X}$. Additionally,
 	\begin{equation*}

### 2021-08-23 manuscripts/laplacian_eigenmaps/sections/introduction.tex
Updated preliminaries section of Laplacian Eigenmaps manuscript.

@@
 \label{sec:introduction}
-Two fundamental goals of statistics, data science, and machine learning are to learn and leverage structure of an unknown distribution, on the basis of samples drawn from said distribution.
+Two fundamental statistical goals are to learn and leverage structure of an unknown distribution, on the basis of samples drawn from said distribution.
 Neighborhood graph Laplacians, and their associated spectra, have proven to be powerful and general tools for both purposes.
@@
 
-By this point there exists a rich literature (e.g.
+By now there exists a rich literature (e.g.
 \citep{koltchinskii2000,belkin07,vonluxburg2008,burago2014,shi2015,singer2017,garciatrillos18,trillos2019, calder2019, cheng2021,dunson2021}) explaining this practical success from a theoretical perspective.
@@
 In contrast, in this paper we develop theory showing that graph spectral methods can effectively \emph{leverage} structure to perform downstream (supervised) learning tasks.
-Our main goal is to show that spectral projection procedures for nonparametric estimation and testing---which use eigenvectors of a neighborhood graph Laplacian as features in linear regression---are statistically optimal.
+Our main goal is to show that spectral projection procedures--which use eigenvectors of a neighborhood graph Laplacian as features in linear regression---are statistically optimal methods for nonparametric estimation and goodness-of-fit testing.
 
-The regression methods we consider are based on \emph{Laplacian eigenmaps}, first introduced by~\cite{belkin03a} \footnote{Technically speaking, as used by~\cite{belkin03a} ``Laplacian eigenmaps'' refers to embedding each design point $X_i$ according to coordinates defined by the first-$K$ graph Laplacian eigenvectors; that is, mapping $X_i \mapsto (v_{i,1},\ldots,v_{i,K})$.
-What we call Laplacian eigenmaps is a two-step supervised method: the first step is the embedding step, and the second step computes the estimate $\wh{f}$ using linear regression.
-Since in this work we only ever deal with the supervised method, and since the choice of embedding step is the far more important of the two steps, we take the liberty of referring to the entire procedure as Laplacian eigenmaps}.
+The regression methods we consider are based on \emph{Laplacian eigenmaps}, first introduced by~\cite{belkin03a} \footnote{Technically speaking, as used by~\cite{belkin03a} ``Laplacian eigenmaps'' refers to a spectral embedding procedure.
+What we call Laplacian eigenmaps is a two-step regression method: the first step is this spectral embedding, and the second step computes the estimate $\wh{f}$ by running ordinary least squares in embedding space.
+Since in this work we only ever consider the embedding within the context of this two-step regression method, we take the liberty of referring to the entire procedure as Laplacian eigenmaps.}.
 Given samples $(X_1,Y_1),\ldots,(X_n,Y_n)$, Laplacian eigenmaps first constructs a neighborhood graph---with vertices at the design points $X_1,\ldots,X_n$ and weighted edges between pairs of sufficiently close points $X_i$ and $X_j$---and computes the corresponding (unweighted) graph Laplacian.
-The Laplacian eigenmaps estimate $\wh{f}$ is simply the projection of the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto the subspace spanned by a subset of the graph Laplacian eigenvectors.
-The empirical norm $\|\wh{f}\|_n^2$ can be used as a statistic to test whether any signal is present: that is, whether the regression function $f_0(x) = \mathbb{E}[Y|X = x]$ is anywhere non-zero.
-(The precise definition of these methods is given in Section~\ref{}.) We study these methods through the lens of a minimax analysis, and give upper bounds on the error of the Laplacian eigenmaps estimator and test which imply that they are minimax optimal over certain smoothness classes.
+The Laplacian eigenmaps estimate $\wh{f}$ is found by projecting the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto a subspace spanned by a subset of the graph Laplacian eigenvectors, and the empirical norm $\|\wh{f}\|_n^2$ can be used as a statistic to test whether any signal is present: that is, whether the regression function $f_0(x) = \mathbb{E}[Y|X = x]$ is anywhere non-zero.
+(The precise definition of these methods is given in Section~\ref{subsec:laplacian_eigenmaps}.) We study these methods through the lens of a minimax analysis.
+Our major contribution is to provide upper bounds on the error of the Laplacian eigenmaps estimator and test which imply that they are minimax optimal over natural smoothness classes.
 
@@
 Roughly speaking, for a domain $\mc{X} \subseteq \Rd$, the order-$s$ (Hilbert-)Sobolev space $H^s(\mc{X})$ consists of those functions which are $s$-times weakly differentiable, with derivatives bounded in $L^2(\mc{X})$ norm.
-Under standard regularity conditions, the minimax estimation and testing rates for Sobolev spaces are well-known: when using mean-squared error as the loss, the estimation rate is $n^{-2s/(2s + d)}$, and the testing rate is $n^{-4s/(4s + d)}$.
-These rates are achieved by various estimators, but classical spectral projection methods\footnote{We use ``classical'' to refer to spectral projection methods using eigenfunctions of a (fixed) Laplacian or Laplace-Beltrami operator, as distinct from Laplacian eigenmaps which uses eigenvectors of a (data-dependent) graph Laplacian matrix.}---which use the eigenfunctions $\psi_1,\psi_2,\ldots$ of a Laplace-Beltrami operator defined on $\mc{X}$---are particularly well-suited for regression over Sobolev spaces.
-This is because Sobolev spaces can be equivalently defined in a spectral manner: they consist of those functions $f$ whose Fourier coefficients $\theta_k := \int_{\mc{X}} f \psi_k$ are bounded in a particular sequence norm.
-For this reason, classical spectral projection methods have excellent statistical properties; see e.g.~\cite{tsybakov2008_book} and \cite{ingster2009} for analysis of estimation and testing risk in the special case where the design distribution is uniform over the unit cube, and Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} in Section~\ref{subsec:minimax_rates_sobolev} of our paper, which give equivalent upper bounds under more general conditions.
-However, these classical spectral methods require a priori knowledge of the design distribution, which is often not available.
+Equivalently, Sobolev spaces can be defined in a spectral manner;  letting $\psi_1,\psi_2,\ldots$ be the eigenfunctions of a Laplace-Beltrami operator defined on $\mc{X}$, Sobolev spaces consist of those functions $f$ whose Fourier coefficients $\theta_k := \int_{\mc{X}} f \psi_k$ are bounded in a particular sequence norm (see Section~\ref{subsec:spectral_projection}).
+The minimax estimation and testing rates for Sobolev spaces are generally speaking well-known---when using mean-squared error as the loss, the estimation rate is $n^{-2s/(2s + d)}$, and the testing rate is $n^{-4s/(4s + d)}$---and are achieved by various estimators.
+However, since Sobolev spaces admit a spectral characterization in terms of $\psi_1,\psi_2,\ldots$, classical spectral projection methods\footnote{We use ``classical'' to refer to spectral projection methods using eigenfunctions of a (fixed) Laplacian or Laplace-Beltrami operator, as distinct from Laplacian eigenmaps which uses eigenvectors of a (data-dependent) graph Laplacian matrix.} using these eigenfunctions are a particularly natural choice for regression over Sobolev spaces, and are in a sense canonical estimators for these function classes.
+% (AG 8/13/21): Last clause makes a strong statement that I personally believe, but may irritate certain readers. Keep or remove?
+% (AG 8/18/21): I do not love the name ``classical spectral projection method''. It doesn't roll off the tongue, and ``classical'' doesn't really convey what the method is. Is there a better name?
 
-On the other hand, Laplacian eigenmaps implicitly learns the structure of the design distribution, and serves as a data-dependent approximation to classical spectral projection that can be used even when the design distribution is unknown.
+Laplacian eigenmaps serves as a data-dependent approximation to these classical spectral projection methods.
+This data dependency can be very useful.
+Computing classical spectral methods requires a priori knowledge of the design distribution, since they are defined with respect to the eigenfunctions of a Laplace-Beltrami operator which itself depends on this distribution.
+Such a priori information is often not available.
+On the other hand, Laplacian eigenmaps implicitly learns structure from the design points $\{X_1,\ldots,X_n\}$, and can thus be used even when the design distribution is unknown.
 Of course, this comes with a tradeoff.
-Intuitively, using an empirical approximation to the ``right'' population-level basis will incur some additional error; the question is whether this additional error is enough to meaningfully degrade the overall statistical properties.
+Intuitively, using an empirical approximation to the ``right'' population-level basis will incur some additional error; the question is whether, notwithstanding this additional error, Laplacian eigenmaps shares the strong statistical properties of classical spectral projection.
 
-We find that in many cases, the answer is no: under appropriate regularity conditions, when correctly tuned Laplacian eigenmaps is minimax optimal over Sobolev spaces.
+We find that in many cases, the answer is yes: under appropriate regularity conditions, when correctly tuned Laplacian eigenmaps is minimax optimal over Sobolev spaces.
 Specifically, our main contributions are as follows:
@@
 In this case, it is known that the minimax rates over Sobolev spaces depend on the intrinsic dimension $m$ of the manifold: explicitly, the estimation rate is $n^{-2s/(2s + m)}$ and the testing rate is $n^{-4s/(4s + m)}$.
-We give upper bounds implying that Laplacian eigenmaps achieves these rates so long as $s \in \{1,2,3\}$, and demonstrating that Laplacian eigenmaps is a~\emph{domain adaptive} estimator.
+We give upper bounds implying that Laplacian eigenmaps achieves these rates so long as $s \in \{1,2,3\}$.
+One can interpret these results as precisely quantifying a way in which Laplacian Eigenmaps leverages structure, by adapting to the intrinsic dimension of the design.
 \end{itemize}
-(In all cases, our bounds also depend optimally on the radius $M$ of the Sobolev ball under consideration.)
+In all cases, our bounds also depend optimally on the radius $M$ of the Sobolev ball under consideration.
 
-To derive these results, we develop an analysis framework reminiscent of that used in the analysis classical spectral projection estimators, but ultimately quite different.
-The chief difference is in the approximation error (bias).
-For our purposes we must establish that, with high probability, the eigenvectors of graph Laplacians effectively approximate continuum Sobolev functions, whereas in the classical case this follows immediately from the aforementioned spectral characterization of Sobolev spaces.
-To tighly upper bound the approximation error of Laplacian eigenmaps, we rely on several recent results regarding concentration of the spectra of graph Laplacians, and prove some additional results which may be of independent interest.
+To derive these results, we develop an analysis framework reminiscent of that used in the analysis of classical spectral projection estimators, but ultimately quite different.
+The chief difference is in the approximation error (bias): for our purposes we must establish that with high probability, the eigenvectors of graph Laplacians (which are data-dependent and hence random objects) effectively approximate continuum Sobolev functions, whereas in the classical case an upper bound on the approximation error follows immediately from the aforementioned spectral characterization of Sobolev spaces.
+To tightly upper bound the approximation error of Laplacian eigenmaps, we rely on several recent results regarding concentration of the spectra of graph Laplacians around their population limits, and prove some additional results which may be of independent interest.
 
@@
 However, for some values of $s$ (number of derivatives) and $d$ (dimension), there do exist gaps between our upper bounds on the error of Laplacian eigenmaps and the minimax rates.
-Although we do not give corresponding lower bounds verifying the tightness of our analysis, we believe these gaps reflect the true behavior of the method rather than some looseness in our analysis; we comment more on this at relevant parts in the text.
+Although we do not give corresponding lower bounds verifying the tightness of our analysis, we believe these gaps reflect the true behavior of the method rather than some looseness in our analysis, and we comment more on this at relevant parts in the text.
 For completeness, we summarize all of our upper bounds---those which match the minimax rates, and those which do not---in Tables~\ref{tbl:estimation_rates} and~\ref{tbl:testing_rates}.
@@
 In each case, rates hold for all $d \in \mathbb{N}$ (under Model~\ref{def:model_flat_euclidean}), and for all $m \in \mathbb{N}, 1 < m < d$ (under Model~\ref{def:model_manifold}).
-Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps estimator is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $n^{-1/2} \lesssim M \lesssim n^{1/d}$.}
+Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps estimator is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal}
 	\label{tbl:estimation_rates}
@@
 Bold font marks minimax optimal rates.
-Rates when $d > 4s$ assume that $f_0 \in L^4(P,M)$.
-Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $n^{-1/2} \lesssim M \lesssim n^{1/d}$.}
+Rates when $d > 4s$ assume that $f_0 \in L^4(\mc{X})$, and depend on $\|f_0\|_{L^4(\mc{X})}$.
+Although we suppress it for simplicity, in all cases when othe Laplacian eigenmaps test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal.}
 	\label{tbl:testing_rates}
@@
 This setting is quite general, because the graph need not be a geometric graph defined on a vertex set which belongs to Euclidean space.
-On the other hand, depending on the data-collection process, it may be unnatural to model the design points as being a priori fixed, and the estimand as being a vector which exhibits a discrete notion of ``smoothness'' over this fixed design.
+On the other hand, depending on the data collection process, it may be unnatural to model the design points as being a priori fixed, and the estimand as being a vector which exhibits a discrete notion of ``smoothness'' over this fixed design.
 Instead, we adopt the \emph{random design} perspective, and seek to estimate a function that we assume exhibits a more classical notion of smoothness.
@@
 In the random design setting the analysis is more challenging, and only a few works have studied the matter.
-\cite{zhou2011} provide an asymptotic analysis of the Laplacian eigenmaps estimator in the semi-supervised setting, as the number of unlabeled samples tends to infinity; in this case the estimator reduces to the classical spectral projection estimator, and there is no extra approximation error due to using data-dependent eigenvectors.
-\cite{lee2016} consider the \emph{diffusion maps} estimator---which uses the eigenvectors of a differently normalized graph Laplacian---in a similar supervised setting to our own, but prove suboptimal rates of convergence relative to the minimax rates.
+\cite{zhou2011} provide an asymptotic analysis of the Laplacian eigenmaps estimator in the semi-supervised setting, as the number of unlabeled samples tends to infinity; in this case the estimator reduces to the classical spectral projection estimator, and there is no extra error incurred by using data-dependent eigenvectors.
+\cite{lee2016} consider the \emph{diffusion maps} estimator---which uses the eigenvectors of a differently normalized graph Laplacian---in a similar supervised setting to our own, but prove suboptimal rates of convergence.
 Finally, \citet{trillos2020,green2021} study a penalized least squares method for estimation, which uses a quadratic form involving the Laplacian as a penalty.
-This method is optimal only for the first-order Sobolev space $H^1(\mc{X})$, and even then only when $d \in \{1,2,3,4\}$, in contrast to the Laplacian eigenmaps estimator which is optimal for all $s$ and $d$; we discuss this striking contrast more in Section~\ref{sec:discussion}.
+This method is optimal only for the first-order Sobolev space $H^1(\mc{X})$, and even then only when $d \in \{1,2,3,4\}$, in contrast to the Laplacian eigenmaps estimator which is optimal for all $s$ and $d$; we discuss this contrast more in Section~\ref{sec:discussion}.
 
@@
 We now outline the structure of the rest of this paper.
-In Section~\ref{sec:setup_main_results}, we formally setup the models and methods we study.
-Propositions~\ref{prop:spectral_series_estimation} and ~\ref{prop:spectral_series_testing}, in Section~\ref{sec:spectral_projection}, show that under rather general conditions on the design distribution, classical spectral projection methods achieve minimax rates of convergence.
-Then in Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps}, \ref{sec:out_of_sample} and and~\ref{sec:manifold_adaptivity}, we give various upper bounds on the error of Laplacian eigenmaps (summarized above), as well as summarizing our analysis at a high-level and providing some key intermediary results.
+In Section~\ref{sec:setup_main_results}, we give our formal modeling assumptions, and precisely define the Laplacian Eigenmaps methods we study.
+Propositions~\ref{prop:spectral_series_estimation} and ~\ref{prop:spectral_series_testing}, in Section~\ref{subsec:spectral_projection}, show that under rather general conditions on the design distribution, classical spectral projection methods achieve minimax rates of convergence over Sobolev classes.
+Then in Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps}, \ref{sec:out_of_sample} and and~\ref{sec:manifold_adaptivity}, we give our main upper bounds on the error of Laplacian eigenmaps (summarized above); these upper bounds certify that under similarly general conditions, Laplacian Eigenmaps is also minimax rate-optimal.
 In Section~\ref{sec:experiments} we examine the empirical behavior of Laplacian eigenmaps, and show that even at moderate sample sizes Laplacian eigenmaps is competitive with classical spectral projection.
@@
 We write $\partial f/\partial r_i$ for the partial derivative of $f$ in the $i$th standard coordinate of $\Rd$, and use the multi-index notation $D^{\alpha}f := \partial^{|\alpha|}f/\partial^{\alpha_1}x_1\ldots\partial^{\alpha_d}x_d$ for multi-indices $\alpha \in \Reals^d$.
+Recall that for a given multi-index $\alpha \in \mathbb{N}^d$, a function $f$ is \emph{$\alpha$-weakly differentiable} if there exists some $h \in L^1(\mc{X})$ such that
+\begin{equation*}
+\int_{\mc{X}} h g = (-1)^{|\alpha|} \int_{\mc{X}} f D^{\alpha}g, \quad \textrm{for every $g \in C_c^{\infty}(\mc{X})$.}
+\end{equation*}
+If such a function $h$ exists, it is the $\alpha$th weak partial derivative of $f$, and denoted by $D^{\alpha}f := h$.
+For functions $f$ which are $|\alpha|$-times classically differentiable, this coincides with the classical definition of derivative, and so we use the same notation for both.
 

### 2021-08-23 manuscripts/laplacian_eigenmaps/sections/preliminaries.tex
Updated preliminaries section of Laplacian Eigenmaps manuscript.

@@
 
-In this section, we precisely define our modeling assumptions and the Laplacian eigenmaps methods we study.
+We begin in Sections~\ref{subsec:regression_laplacian_eigenmaps}-\ref{subsec:laplacian_eigenmaps} by precisely defining the models (random design points, Sobolev-smooth regression functions) and methods (Laplacian eigenmaps) under consideration.
+Then in Section~\ref{subsec:spectral_projection}, we connect Sobolev spaces to Laplacian eigenmaps using the spectrum of a density-dependent Laplace-Beltrami operator, and show in Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} that projection methods which use the eigenfunctions of this operator are statistically optimal.
 
 \subsection{Nonparametric regression over Sobolev spaces}
-\label{sec:regression_laplacian_eigenmaps}
+\label{subsec:regression_laplacian_eigenmaps}
 
-We will operate in the usual setting of nonparametric regression with random design, in which we observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$.
-The design points $X_1,\ldots,X_n$ are sampled from a distribution $P$ with support $\mc{X} \subseteq \Rd$, and the responses follow the signal plus noise model
+We will always operate in the usual setting of nonparametric regression with random design.
+We observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$; the design points $X_1,\ldots,X_n$ are sampled from a distribution $P$ with support $\mc{X} \subseteq \Rd$, and the responses follow the signal plus noise model
 \begin{equation}
@@
 with regression function $f_0: \mc{X} \to \Reals$, and $w_i \sim N(0,1)$ independent Gaussian noise. 
-% (AG): Removed the comment regarding noise with variance = \sigma^2.
+% (AG 8/4/21): Removed the comment regarding noise with variance = \sigma^2.
 
@@
 
+\paragraph{Flat Euclidean model.}
+In Definitions~\ref{def:model_flat_euclidean}-\ref{def:zero_trace_sobolev_space}, we collect the assumptions we make when working under the flat Euclidean model.
+We begin by giving some regularity conditions on the design.
+
 \begin{definition}[Flat Euclidean model]
 	\label{def:model_flat_euclidean}
-	The data $(X_1,Y_1),\ldots,(X_n,Y_n)$ are sampled according to~\eqref{eqn:model}.
 	 The support $\mc{X}$ of the design distribution $P$ is an open, connected, and bounded subset of $\Rd$, with Lipschitz boundary.
@@
 \end{definition}
+% (AG 8/13/21): I got rid of the sentence ``The data are sampled according to (1).'' It seemed repetitive. 
+At various points we will also assume that the density $p \in C^k(\mc{X})$ for some $k \in \mathbb{N}$; when we make such assumptions we will state them explicitly.
 
-In the following, we recall that the injectivity radius of an $m$-dimensional Riemannian manifold $\mc{X}$ is the maximum value of $\delta$ such that the exponential map $\exp_x: B_m(0,\delta) \subset T_x(\mc{X}) \to B_{\mc{X}}(x,\delta) \subset \mc{X}$ is a diffeomorphism for all $x \in \mc{X}$.
-\begin{definition}[Manifold model]
-	\label{def:model_manifold}
-	The data $(X_1,Y_1),\ldots,(X_n,Y_n)$ are sampled according to~\eqref{eqn:model}. 
-	The support $\mc{X}$ of the design distribution $P$ is a closed, connected, smooth and boundaryless Riemannanian manifold embedded in $\Rd$, of intrinsic dimension $1 \leq m < d$.
-The injectivity radius of $\mc{X}$ is lower bounded by a positive constant $i_0 > 0$.
-The design distribution $P$ admits a Lipschitz density $p$ with respect to the volume form $d\mu$ induced by the Riemannian structure of $\mc{X}$, which is bounded away from $0$ and $\infty$,
-	\begin{equation*}
-	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty, \quad \textrm{for all $x \in \mc{X}$.}
-	\end{equation*}
-\end{definition}
-
-Finally, at various points we will have to assume that the density $p$ also satisfies different types of higher-order regularity, beyond Lipschitz continuity.
-When such assumptions are necessary we will state them explicitly.
-
-\subsection{Laplacian eigenmaps}
-We now formally define the estimator and test statistic we study.
-Both are derived from eigenvectors of a graph Laplacian.
-For a positive, symmetric kernel $\eta: [0,\infty) \to [0,\infty)$, and a radius parameter $\varepsilon > 0$, let $G = ([n],W)$ be the neighborhood graph formed over the design points $\{X_1,\ldots,X_n\}$, with a weighted edge $W_{ij} = \eta(\|X_i - X_j\|/\varepsilon)$ between vertices $i$ and $j$.
-Then the
-\emph{neighborhood graph Laplacian} $L_{n,\varepsilon}: \Reals^n \to \Reals$ is defined by its action on vectors $u \in \Reals^n$ as
-\begin{equation}
-\label{eqn:neighborhood_graph_laplacian}
-\bigl(L_{n,\varepsilon}u\bigr)_i := \frac{1}{n\varepsilon^{2 + \mathrm{dim}(\mc{X})}} \sum_{j = 1}^{n} \bigl(u_i - u_j\bigr) \eta\biggl(\frac{\|X_i - X_j\|}{\varepsilon}\biggr).
-\end{equation}
-(Here $\mathrm{dim}(\mc{X})$ stands for the dimension of $\mc{X}$.
-It is equal to $d$ under the assumptions of Model~\ref{def:model_flat_euclidean}, and equal to $m$ under the assumptions of Model~\ref{def:model_manifold}.
-The pre-factor $(n\varepsilon^{2 + \mathrm{dim}(\mc{X})})^{-1}$ is purely for convenience in taking limits as $n \to \infty, \varepsilon \to 0$).
-Written in standard coordinates we have $(n\varepsilon^{\dim(\mc{X}) + 2}) \cdot L_{n,\varepsilon} = D - W$, where $D \in \Reals^{n \times n}$ is the diagonal degree matrix, $D_{ii} = \sum_{i = 1}^{n} W_{ij}$.
-
-The graph Laplacian is a positive semi-definite matrix, and admits the eigendecomposition $L_{n,\varepsilon} = \sum_{k = 1}^{n} \lambda_k v_k v_k^{\top}$, where for each $k = 1,\ldots,n$ the eigenvalue-eigenvector pair $(\lambda_k,v_k)$ satisfies
-\begin{equation*}
-L_{n,\varepsilon}v_k = \lambda_k v_k, \quad \|v_k\|_2^2 = 1.
-\end{equation*}
-We will assume without loss of generality that each eigenvalue $\lambda$ of $L_{n,\varepsilon}$ has algebraic multiplicity $1$, and so we can index the eigenpairs $(\lambda_1,v_1),\ldots,(\lambda_n,v_n)$ in ascending order of eigenvalue, $0 = \lambda_1 < \ldots < \lambda_n$.
-
-The Laplacian eigenmaps estimator $\wh{f}$ simply projects the response vector ${\bf Y}$ onto the first $K$ eigenvectors of $L_{n,\varepsilon}$: letting $V_K \in \Reals^{n \times K}$ be the matrix with columns $v_1,\ldots,v_K$, we have that
-\begin{equation}
-\label{eqn:laplacian_eigenmaps_estimator}
-\wh{f} := \sum_{k = 1}^{K} \dotp{{\bf Y}}{v_k}_{n} v_k = \frac{1}{n} V_K V_K^{\top} {\bf Y}.
-\end{equation} 
-Thus $\wh{f}$ is equivalently a vector in $\Reals^n$, or a function in $L^2(P_n)$. If $\wh{f}$ is a reasonable estimate of $f_0$, then the Laplacian eigenmaps test statistic
-\begin{equation}
-\label{eqn:laplacian_eigenmaps_test}
-\wh{T} := \|\wh{f}\|_n^2
-\end{equation}
-is in turn a reasonable estimate of $\|f_0\|_{P}^2$, and can be used to distinguish whether or not $f_0 = 0$.
-
-% AG 8/12/21: Commented out due to overlap with footnote in new introduction.
-% It may be helpful to comment briefly on the term ``Laplacian eigenmaps'', which we use a bit differently than is typical in the literature.
-Laplacian eigenmaps typically refers to an algorithm for embedding, which maps each design point $X_1,\ldots,X_n$ to $\Reals^K$ according to $X_i \mapsto (v_{1,i}, \ldots, v_{K,i})$.
-Viewing this embedding as a feature map, we can then interpret the estimator $\wh{f}$ as the least-squares solution to a linear regression problem with responses $Y_1,\ldots,Y_n$ and features $v_1,\ldots,v_K$.
-Often, the Laplacian eigenmaps embedding is viewed as a tool for dimensionality reduction, wherein it is implicitly assumed that $K$ is much smaller than $d$.
-We will neither explicitly nor implicitly take $K < d$; indeed, the embedding perspective is not particularly illuminating in what follows, and we do not henceforth make reference to it.
-Instead, we use ``Laplacian eigenmaps'' to directly refer to the estimator $\wh{f}$ or test statistic $\wh{T}$.
-
-\subsection{Sobolev Classes}
-\label{sec:sobolev}
-We now review the definition of Sobolev classes, dividing our discussion into two cases, the flat Euclidean case (Model~\ref{def:model_flat_euclidean}) and the manifold case (Model~\ref{def:model_manifold}).
-
-\paragraph{Flat Euclidean case.}
-Take $\mc{X}$ to be a open, connected and bounded set with Lipschitz boundary, as in Model~\ref{def:model_flat_euclidean}.
-Recall that for a given multi-index $\alpha \in \mathbb{N}^d$, a function $f$ is \emph{$\alpha$-weakly differentiable} if there exists some $h \in L^1(\mc{X})$ such that
-\begin{equation*}
-\int_{\mc{X}} h g = (-1)^{|\alpha|} \int_{\mc{X}} f D^{\alpha}g, \quad \textrm{for every $g \in C_c^{\infty}(\mc{X})$.}
-\end{equation*}
-If such a function $h$ exists, it is the $\alpha$th weak partial derivative of $f$, and denoted by $D^{\alpha}f := h$.
-For functions $f$ which are $|\alpha|$-times differentiable, this coincides with the classical definition of derivative.
-
+We model the regression function as belonging to an order-$s$ Sobolev space, and being bounded in Sobolev norm.
 \begin{definition}[Sobolev space on a flat Euclidean domain]
@@
 \end{definition}
-We note that $H^s(\mc{X})$ is the completion of $C^{\infty}(\mc{X})$ with respect to the $\|\cdot\|_{H^s(\mc{X})}$ norm, so that $C^{\infty}(\mc{X})$ is dense in $H^s(\mc{X})$.
+Finally when $s > 1$ we will assume that $f_0$ satisfies a zero-trace boundary condition.
+Recall that $H^s(\mc{X})$ can alternatively be defined as the completion of $C^{\infty}(\mc{X})$ in the Sobolev norm $\|\cdot\|_{H^s(\mc{X})}$.
+The zero-trace Sobolev spaces are defined in a similar fashion, as the completion of $C_c^{\infty}(\mc{X})$ in the same norm.
 
-\paragraph{Manifold case.}
+\begin{definition}[Zero-trace Sobolev space]
+	\label{def:zero_trace_sobolev_space}
+	A function $f \in H^s(\mc{X})$ belongs to the zero-trace Sobolev space $H_0^s(\mc{X})$ if there exists a sequence $f_1,f_2,\ldots$ of functions in $C_c^{\infty}(\mc{X})$ such that
+	\begin{equation*}
+	\lim_{k \to \infty}\|f_k - f\|_{H^s(\mc{X})} = 0.
+	\end{equation*}
+	The normed ball $H_0^{s}(\mc{X};M) := H_0^{s}(\mc{X}) \cap H^{s}(\mc{X};M)$.
+\end{definition}
+Boundary conditions plays an important role in the analysis of spectral methods, as we explain further in Section~\ref{subsec:spectral_projection}.
+For now, we limit ourselves to pointing out that for functions $f \in C^\infty(\mc{X})$, the zero-trace condition implies that $\partial^{k}f/\partial{\bf n}^k(x) = 0$ for each $k = 0,\ldots,s - 1$, and for all $x \in \partial\mc{X}$.
+(Here $\partial/(\partial {\bf n})$ is the partial derivative operator in the direction of the normal vector $\mathbf{n}$.)
+\paragraph{Manifold model.}
+As in the flat Euclidean case, we start with some regularity conditions on the design.
+\begin{definition}[Manifold model]
+	\label{def:model_manifold}
+	The support $\mc{X}$ of the design distribution $P$ is a closed, connected, and smooth Riemannian manifold (without boundary) embedded in $\Rd$, of intrinsic dimension $1 \leq m < d$, and with a positive injectivity radius $\mathrm{inj}(\mc{X}) > 0$.
+The design distribution $P$ admits a Lipschitz density $p$ with respect to the volume form $d\mu$ induced by the Riemannian structure of $\mc{X}$, which is bounded away from $0$ and $\infty$,
+	\begin{equation*}
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty, \quad \textrm{for all $x \in \mc{X}$.}
+	\end{equation*}
+\end{definition}
+In the above, recall that the injectivity radius $\mathrm{inj}(\mc{X})$ of an $m$-dimensional Riemannian manifold $\mc{X}$ is the maximum $\delta > 0$ such that the exponential map $\exp_x: B_m(0,\delta) \subset T_x(\mc{X}) \to B_{\mc{X}}(x,\delta) \subset \mc{X}$ is a diffeomorphism for all $x \in \mc{X}$.
 
-There are several equivalent ways to define Sobolev spaces on a compact, smooth, $m$-dimensional Riemannian manifold embedded in $\Reals^d$.
+There are several equivalent ways to define Sobolev spaces on smooth Riemannian manifolds.
 We will stick with a definition that parallels our setup in the flat Euclidean setting as much as possible.
@@
 The partial derivatives $D^{\alpha}f$ will depend on the choice of local coordinates, and so will the resulting Sobolev norm~$\|f\|_{H^s(\mc{X})}$.
-However, the important point is that regardless of the choice of local coordinates the resulting norms will be equivalent\footnote{Recall that norms $\|\cdot\|_1$ and $\|\cdot\|_2$ on a space $\mc{F}$ are said to be equivalent if there exist constants $c$ and $C$ such that
+However, for our purposes the important point is that regardless of the choice of local coordinates the resulting norms will be equivalent\footnote{Recall that norms $\|\cdot\|_1$ and $\|\cdot\|_2$ on a space $\mc{F}$ are said to be equivalent if there exist constants $c$ and $C$ such that
 	\begin{equation*}
@@
 and so the ultimate Sobolev space $H^s(\mc{X})$ is independent of the choice of local coordinates.
+For more information regarding manifolds and Sobolev spaces defined thereupon, see~\cite{lee2013} and~\cite{hebey1996}.
 
-\paragraph{Boundary conditions.}
-In the flat Euclidean model (Model~\ref{def:model_flat_euclidean}), in order to show that Laplacian eigenmaps is optimal over $H^{s}(\mc{X})$ for $s > 1$ we will need to assume that the regression function $f_0$ satisfies some boundary conditions.
-In particular, we will assume that $f_0$ is zero-trace.
-\begin{definition}[Zero-trace Sobolev space]
-	\label{def:zero_trace_sobolev_space}
-	The \emph{order-s zero-trace Sobolev space} $H_0^s(\mc{X})$ is the closure of $C_c^\infty(\mc{X})$ with respect to $\|\cdot\|_{H^s(\mc{X})}$ norm.
-That is, $f \in H_0^s(\mc{X})$ if $f \in H^s(\mc{X})$ and additionally there exists a sequence $f_1,f_2,\ldots$ of functions in $C_c^{\infty}(\mc{X})$ such that
+\subsection{Laplacian eigenmaps}
+\label{subsec:laplacian_eigenmaps}
+We now formally define the estimator and test statistic we study.
+Both are derived from eigenvectors of a graph Laplacian.
+For a positive, symmetric kernel $\eta: [0,\infty) \to [0,\infty)$, and a radius parameter $\varepsilon > 0$, let $G = ([n],W)$ be the neighborhood graph formed over the design points $\{X_1,\ldots,X_n\}$, with a weighted edge $W_{ij} = \eta(\|X_i - X_j\|/\varepsilon)$ between vertices $i$ and $j$.
+Then the
+\emph{neighborhood graph Laplacian} $L_{n,\varepsilon}: \Reals^n \to \Reals$ is defined by its action on vectors $u \in \Reals^n$ as
+\begin{equation}
+\label{eqn:neighborhood_graph_laplacian}
+\bigl(L_{n,\varepsilon}u\bigr)_i := \frac{1}{n\varepsilon^{2 + \mathrm{dim}(\mc{X})}} \sum_{j = 1}^{n} \bigl(u_i - u_j\bigr) \eta\biggl(\frac{\|X_i - X_j\|}{\varepsilon}\biggr).
+\end{equation}
+(Here $\mathrm{dim}(\mc{X})$ stands for the dimension of $\mc{X}$.
+It is equal to $d$ under the assumptions of Model~\ref{def:model_flat_euclidean}, and equal to $m$ under the assumptions of Model~\ref{def:model_manifold}.
+The pre-factor $(n\varepsilon^{2 + \mathrm{dim}(\mc{X})})^{-1}$ ensures non-degenerate stable limits as $n \to \infty, \varepsilon \to 0$).
+Written in standard coordinates we have $(n\varepsilon^{\dim(\mc{X}) + 2}) \cdot L_{n,\varepsilon} = D - W$, where $D \in \Reals^{n \times n}$ is the diagonal degree matrix, $D_{ii} = \sum_{i = 1}^{n} W_{ij}$.
+
+The graph Laplacian is a positive semi-definite matrix, and admits the eigendecomposition $L_{n,\varepsilon} = \sum_{k = 1}^{n} \lambda_k v_k v_k^{\top}$, where for each $k = 1,\ldots,n$ the eigenvalue-eigenvector pair $(\lambda_k,v_k)$ satisfies
 \begin{equation*}
-	\lim_{k \to \infty}\|f_k - f\|_{H^s(\mc{X})} = 0.
+L_{n,\varepsilon}v_k = \lambda_k v_k, \quad \|v_k\|_2^2 = 1.
 \end{equation*}
-	The normed ball $H_0^{s}(\mc{X};M) := H_0^{s}(\mc{X}) \cap H^{s}(\mc{X};M)$.
-\end{definition}
-The zero-trace condition can be made more concrete when $f \in C^\infty(\mc{X})$, since we can then speak of the pointwise behavior of $f$ and its derivatives.
-Letting $\partial/(\partial {\bf n})$ be the partial derivative operator in the direction of the vector $\mathbf{n}$ normal to the boundary of $\mc{X}$, then the zero-trace condition implies that $\partial^{k}f/\partial{\bf n}^k(x) = 0$ for each $k = 0,\ldots,s - 1$, and for all $x \in \partial\mc{X}$.
+We will assume without loss of generality that each eigenvalue $\lambda$ of $L_{n,\varepsilon}$ has algebraic multiplicity $1$, and so we can index the eigenpairs $(\lambda_1,v_1),\ldots,(\lambda_n,v_n)$ in ascending order of eigenvalue, $0 = \lambda_1 < \ldots < \lambda_n$.
 
-% (SB via AG) Move following chunk to appendix, and replace with one sentence explaining that when studying spectral projection methods, it is standard to impose boundary conditions.
-We now explain why Laplacian eigenmaps should be optimal only when $f_0$ satisfies certain boundary conditions.
-Let $(\lambda_1(\Delta_P),\psi_1),(\lambda_2(\Delta_P),\psi_2),\ldots$ be the solutions to the weighted Laplacian eigenvector equation with Neumann boundary conditions,
+The Laplacian eigenmaps estimator $\wh{f}$ simply projects the response vector ${\bf Y}$ onto the first $K$ eigenvectors of $L_{n,\varepsilon}$: letting $V_K \in \Reals^{n \times K}$ be the matrix with columns $v_1,\ldots,v_K$, we have that
 \begin{equation}
-\label{eqn:laplacian_eigenfunction_neumann}
-\Delta_P\psi_k = \lambda_{k}(\Delta_P) \psi_k, \quad \frac{\partial}{\partial {\bf n}}\psi_k = 0~~\textrm{on $\partial \mc{X}$}.
+\label{eqn:laplacian_eigenmaps_estimator}
+\wh{f} := \sum_{k = 1}^{K} \dotp{{\bf Y}}{v_k}_{2} v_k = V_K V_K^{\top} {\bf Y}.
 \end{equation} 
-As we have already alluded to, it is known~\citep{garciatrillos18} that each graph Laplacian eigenpair ($\lambda_k$,$v_k$) converges to a corresponding solution ($\lambda_k(\Delta_P)$, $\psi_k$) of~\eqref{eqn:laplacian_eigenfunction_neumann}.
-Thus it is relevant to consider which Sobolev functions $f \in H^s(M)$ we can reconstruct using the eigenfunctions $\psi_1,\psi_2,\ldots$.
-To that end, we introduce the spectrally defined Sobolev space
+If $\wh{f}$ is a reasonable estimate of $f_0$, then the Laplacian eigenmaps test statistic
 \begin{equation}
-\label{eqn:spectral_sobolev_space}
-\mc{H}^s(\mc{X}) := \biggl\{f \in L^2(\mc{X}): \sum_{k = 1}^{\infty} \bigl[\dotp{f}{\psi_k}_P\bigr]^2 \cdot\bigl[\lambda_k(\Delta_P)\bigr]^s < \infty \biggr\}.
+\label{eqn:laplacian_eigenmaps_test}
+\wh{T} := \|\wh{f}\|_n^2 = \frac{1}{n} {\bf Y}^{\top} V_K V_K^{\top} {\bf Y}
 \end{equation}
-Under the conditions $p \in C^{\infty}(\mc{X})$ and $\partial\mc{X} \in C^{1,1}$, \cite{dunlop2020} show the strict inclusion $\mc{H}^{2s}(\mc{X}) \subset H^{2s}(\mc{X})$.
-More precisely, they show that for any $s > 0$,
+is in turn a reasonable estimate of $\|f_0\|_{P}^2$, and can be used in the \emph{signal detection} problem to distinguish whether or not $f_0 = 0$.
+
+% AG 8/12/21: Commented out due to overlap with footnote in new introduction.
+% It may be helpful to comment briefly on the term ``Laplacian eigenmaps'', which we use a bit differently than is typical in the literature.
+Laplacian eigenmaps typically refers to an algorithm for embedding, which maps each design point $X_1,\ldots,X_n$ to $\Reals^K$ according to $X_i \mapsto (v_{1,i}, \ldots, v_{K,i})$.
+Viewing this embedding as a feature map, we can then interpret the estimator $\wh{f}$ as the least-squares solution to a linear regression problem with responses $Y_1,\ldots,Y_n$ and features $v_1,\ldots,v_K$.
+Often, the Laplacian eigenmaps embedding is viewed as a tool for dimensionality reduction, wherein it is implicitly assumed that $K$ is much smaller than $d$.
+We will neither explicitly nor implicitly take $K < d$; indeed, the embedding perspective is not particularly illuminating in what follows, and we do not henceforth make reference to it.
+Instead, we use ``Laplacian eigenmaps'' to directly refer to the estimator $\wh{f}$ or test statistic $\wh{T}$.
+
+\subsection{Classical Spectral Projection over Sobolev spaces}
+\label{subsec:spectral_projection}
+Laplacian Eigenmaps serves as a data-dependent alternative to more classical spectral projection methods constructed out of the eigenfunctions of a Laplace-Beltrami operator (or, in the special case where $\mc{X} = [0,1]^d$, out of a Fourier basis).
+We focus on the following density weighted Laplace-Beltrami operator $\Delta_P:C^2(\mc{X}) \to L^2(\mc{X})$,
 \begin{equation}
-\label{eqn:spectral_sobolev_space_equivalence}
-\mc{H}^{2s}(\mc{X}) = \biggl\{f \in H^{2s}(\mc{X}): \frac{\partial \Delta_P^rf}{\partial {\bf n}} = 0~\textrm{on}~\partial\mc{X},~~\textrm{for all $0 \leq r \leq s - 1$} \biggr\},
+\label{eqn:laplace_beltrami}
+\Delta_Pf := -\frac{1}{p} \mathrm{div}(p^2\nabla f),
 \end{equation}
-and for any $s \geq 0$, $\mc{H}^{2s + 1}(\mc{X}) = \mc{H}^{2s}(\mc{X}) \cap H^{2s + 1}(\mc{X})$.
-If $P$ is uniform on $\mc{X}$, then~\eqref{eqn:spectral_sobolev_space_equivalence} means that a Sobolev function $f \in H^{s}(\mc{X})$ additionally belongs to $\mc{H}^{s}(\mc{X})$ only if all its odd lower order derivatives vanish at the boundary $\partial\mc{X}$.
+with eigenvalue/eigenfunction pairs $(\lambda_1(\Delta_P),\psi_1),(\lambda_2(\Delta_P),\psi_2),\ldots$ defined as solutions to
+\begin{equation}
+\label{eqn:laplace_beltrami_eigenproblem}
+\Delta_P\psi = \lambda(\Delta_P) \psi, \quad \frac{\partial}{\partial{\bf n}}\psi = 0~~\textrm{on $\partial \mc{X}$,}
+\end{equation}
+and sorted as usual in ascending order of eigenvalue.\footnote{For formal justification of why~\eqref{eqn:laplace_beltrami_eigenproblem} has discrete spectrum under either Model~\ref{def:model_flat_euclidean} or~\ref{def:model_manifold}, see~\cite{garciatrillos18,trillos2019}.} In this case, classical spectral projection refers to the estimator and test statistic
+\begin{equation}
+\label{eqn:classical_spectral_projection}
+\wt{f}(x) := \frac{1}{n}\sum_{k = 1}^{K} \dotp{{\bf Y}}{\psi_k}_{n} \psi_k(x),\quad\textrm{and}\quad \wt{T} := \|\wt{f}\|_n^2.
+\end{equation}
+The statistical behavior of these more classical methods is relevant because the spectrum of $\Delta_P$ serves as the population-level limit of graph Laplacian spectra: for any fixed $k \in \mathrm{N}$, as $n \to \infty$
+\begin{equation}
+\lambda_k \to \lambda_k(\Delta_P), \quad\textrm{and}\quad \Bigl\|\frac{1}{n}v_k - \psi_k\Bigr\|_n^2 \to 0,
+\end{equation}
+in either the flat Euclidean (Model~\ref{def:model_flat_euclidean}) or manifold (Model~\ref{def:model_manifold}) setups~\citep{garciatrillos18,trillos2019}.
+It follows that for any fixed $K \in \mathbb{N}$, the Laplacian eigenmaps estimator converges to the $\wh{f} \to \wt{f}$, and likewise $\wh{T} \to \wt{T}$, justifying why we view Laplacian Eigenmaps as a data-dependent alternative to classical spectral projection.
 
-The bottom line is that the eigenvectors $v_k$ of the graph Laplacian accurately approximate only those functions $f \in H^{s}(\mc{X})$ which satisfy some additional boundary conditions.
-Although the zero-trace boundary condition is more restrictive than~\eqref{eqn:spectral_sobolev_space_equivalence}---since it also requires that derivatives of even-order be equal to $0$---the point is that some kind of boundary condition will be necessary in order to obtain optimal rates.
-We will stick to the zero-trace condition, since it greatly eases some of the steps in our proofs.
-% (SB via AG) Move above chun to appendix.
+On the other hand, the eigenvectors and eigenfunctions of a (possibly density-weighted) Laplace-Beltrami operator such as $\Delta_P$ can also be used to give a spectral definition of Sobolev spaces.
+For instance, consider the ellipsoid
+\begin{equation}
+\label{eqn:sobolev_ellipsoid}
+\mc{H}^{s}(\mc{X}) := \Bigl\{\sum_{k = 1}^{\infty} a_k \psi_k \in L^2(\mc{X}):  \sum_{k = 1}^{\infty} a_k^2 \lambda_{k}(\Delta_P) \leq M^2 \Bigr\},
+\end{equation}
+with corresponding norm $\|\sum_{k = 1}^{\infty} a_k \psi_k\|_{\mc{H}^s(\mc{X})}^2 = \sum_{k = 1}^{\infty} a_k^2 \lambda_{k}(\Delta_P)$.
+Under appropriate regularity conditions $\mc{H}^s(\mc{X})$ consists of functions $f \in H^s(\mc{X})$ which also satisfy some additional boundary conditions.
+For instance, assuming Model~\ref{def:model_flat_euclidean}, $p \in C^{\infty}(\mc{X})$ and $\partial \mc{X} \in C^{1,1}$, \citet{dunlop2020} show that the ellipsoid $\mc{H}^{2s}(\mc{X})$ satisfies
+\begin{equation}
+\label{eqn:sobolev_ellipsoid_to_sobolev_ball}
+\mc{H}^{2s}(\mc{X}) = 
+\biggl\{f \in H^{2s}(\mc{X}): \frac{\partial \Delta_P^rf}{\partial {\bf n}} = 0~\textrm{on}~\partial\mc{X},~~\textrm{for all $0 \leq r \leq s - 1$} \biggr\},
+\end{equation}
+for any $s > 0$, and likewise $\mc{H}^{2s + 1}(\mc{X}) = \mc{H}^{2s}(\mc{X}) \cap H^{2s + 1}(\mc{X})$ for any $s \geq 0$; additionally, the norms $\|\cdot\|_{\mc{H}^s(\mc{X})}$ and $\|\cdot\|_{H^s(\mc{X})}$ are equivalent.
 
-On the other hand, in the manifold model (Model~\ref{def:model_manifold}) the domain $\mc{X}$ is without boundary---precisely, every point $x \in \mc{X}$ has a neighborhood that is homeomorphic to an open set in $\Reals^m$, for instance the open ball $B(x,\delta)$ for any $\delta$ smaller than the injectivity radius $i_0$---and so boundary conditions are irrelevant.
+The spectral formulation of Sobolev spaces suggest that classical spectral projection methods---and in turn Laplacian Eigenmaps---are natural choices for regression over such function classes.
+Further, given the similarities between Laplacian Eigenmaps and classical spectral projection methods, the statistical properties of the latter---which are better understood and easier to derive---should intuitively shed some light on the properties of the former.
+To this end, we now give a pair of results showing that classical spectral methods are statistically optimal for nonparametric regression over Sobolev classes with random design, \emph{so long as the design distribution is known}.
+These results are in line with previously known upper bounds on the estimation and testing error of classical spectral methods~\citep{tsybakov08,ingster2009}, but hold under less stringent conditions on the design distribution.
+This latter point is important, since our interest in Laplacian Eigenmaps stems in part from its adaptivity to general and unknown designs.
 
-\subsection{Minimax Rates and Spectral Projection Methods}
-\label{subsec:minimax_rates_sobolev}
-We now review the minimax estimation and goodness-of-fit testing rates over Sobolev balls.
-We will pay special attention to certain classical spectral projection methods which achieve these rates.
-This is because, as we have already discussed, classical spectral projection methods are very related to Laplacian eigenmaps.
-
-\paragraph{Estimation rates over Sobolev balls.}
-In the estimation problem, we ask for an estimator---formally, a Borel measurable function $\wh{f}$ that maps from $\mc{X}$ to $\Reals$---which is close to the regression function $f_0$ with respect to the squared norm $\|\wh{f} - f_0\|_{P}^2$.
-Under Model~\ref{def:model_flat_euclidean}, the minimax estimation rate over the order-$s$ Sobolev ball is
-\begin{equation}
-\label{eqn:sobolev_space_minimax_estimation_rate}
-\inf_{\wh{f}} \sup_{f_0} \Ebb\|\wh{f} - f_0\|_P^2 \asymp M^2(M^2n)^{-2s/(2s + d)};
-\end{equation}
-here the infimum is taken over all estimators $\wh{f}$, and the supremum over all $f_0 \in H^1(\mc{X};M)$ (first-order case) or $f_0 \in H_0^s(\mc{X};M)$ (higher-order case), and we assume $n^{-1/2} \lesssim M$.
-% (SB via AG) Supremum also taken over all densities p? Do the densities have to satisfy certain conditions?
-
-% (SB via AG) 
-% 	1. Not clear that $\wt{f}$ is an estimator in Model 2.1.
-%   2. My description of the analysis sells the proposition short.
-%   3. Replace the proof description with the proof itself.
-The lower bound in~\eqref{eqn:sobolev_space_minimax_estimation_rate} is due to~\citep{stone1980} (at least for the case of $M$ constant, as is most typically considered).
-The upper bound can be certified by a particular spectral projection estimator $\wt{f}$,
-\begin{equation}
-\label{eqn:spectral_series_estimator}
-\wt{f} := \sum_{k = 1}^{K} \dotp{Y}{\psi_k}_n \psi_k,
-\end{equation}
-where $\psi_k$ are the eigenfunctions of $\Delta_P$ defined in~\eqref{eqn:laplacian_eigenfunction_neumann}.
-The optimality of spectral projection estimators over Sobolev type spaces is generally well-understood.
-For instance, see \cite{tsybakov08,johnstone2011,gine16}, who work in the Gaussian sequence model and show that analogous estimators are optimal over Sobolev ellipsoids.
-However, we have not found an analysis of the specific estimator~\eqref{eqn:spectral_series_estimator} under Model~\ref{def:model_flat_euclidean}, and so for completeness we state the result in the following proposition.
+\paragraph{Estimation.}
+We begin with an upper bound on the $L^2(P)$ risk of $\wt{f}$.
 \begin{proposition}
 	\label{prop:spectral_series_estimation}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $\partial \mc{X} \in C^{1,1}$, that $p \in C^{\infty}(\mc{X})$, and that $f_0 \in H^1(\mc{X};M)$ (first-order case) or $f_0 \in H_0^s(\mc{X};M)$ for some $s > 1$ (higher-order case).
-Then there exists a constant $C$ which does not depend on $f_0,M$ or $n$ such that the following statement holds: if the spectral projection estimator $\wt{f}$ is computed with parameter $K = \floor{M^2n}^{d/(2s + d)}$, then
+	Suppose data is observed according to Model~\ref{def:model_flat_euclidean}, and additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$, $f_0 \in \mc{H}^{s}(\mc{X};M)$ and $\|f_0\|_P^2 \leq 1$.
+Then there exists a constant $C$ which does not depend on $f_0,M$ or $n$ such that the following statement holds: if the spectral projection estimator $\wt{f}$ is computed with parameter $K = \floor{M^2n}^{d/(2s + d)} \vee 1$, then
+	\begin{equation}
+	\label{eqn:spectral_series_estimation}
+	\Ebb\bigl[\|\wt{f} - f_0\|_P^2\bigr] \leq C \min\bigl\{M^2(M^2n)^{-2s/(2s + d)}, M^2\bigr\}.
+	\end{equation}
+\end{proposition}
+When the Sobolev ball radius $M \asymp 1$, the upper bound in~\eqref{eqn:spectral_series_estimation} is on the order of $n^{-2s/(2s + d)}$, which matches the standard minimax estimation rate for Sobolev classes (see~\cite{wasserman2006,tsybakov08} and references therein).
+
+% (AG 8/19/21): I would like to strengthen this statement to not require $M \asymp 1$, but I do not know where I can find the rate M^2(M^2n)^{-2s/(2s + d)} for all s and d combinations.
+In fact, I don't even know of a paper that states the M^2(M^2n)^{-2s/(2s + d)} rate for L^2(P) risk in random design.
+
+We now give the proof of Proposition~\ref{prop:spectral_series_estimation}; the structure of the analysis, which is fairly classical and straightforward, is similar to the high-level strategy we use to analyze Laplacian Eigenmaps (see Section~\ref{subsec:analysis}).
+
+\paragraph{Proof of Proposition~\ref{prop:spectral_series_estimation}.}
+We decompose risk into squared bias and variance,
+\begin{equation}
+\label{pf:spectral_series_estimation_0}
+\Ebb \|\wt{f} - f_0\|_P^2 = \Ebb\| \Ebb[\wt{f}]  - f_0\|_P^2 + \Ebb\| \wt{f} - \Ebb[\wt{f}]\|_P^2.
+\end{equation}
+Since the eigenfunctions $\{\psi_k\}$ form an orthonormal basis of $L^2(P)$, and $f_0 \in H_0^{s}(\mc{X}) \subseteq L^2(P)$, we can write the squared bias in terms of squared Fourier coefficients of $f_0$, leading to the following upper bound,
 \begin{equation*}
-	\Ebb\bigl[\|\wt{f} - f_0\|_P^2\bigr] \leq C \min\bigl\{M^2(M^2n)^{-2s/(2s + d)}, M^2\bigr\}
+\|f_0 - \Ebb{\wt{f}}\|_P^2 = \sum_{k = K + 1}^{\infty}  \dotp{f_0}{\psi_k}^2 \leq  \frac{1}{\lambda_{K + 1}(\Delta_P)^s} \sum_{k = K + 1}^{\infty} \lambda_{k + 1}(\Delta_P)^s \dotp{f_0}{\psi_k}^2 \leq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}}{\bigl[\lambda_{K + 1}(\Delta_P)\bigr]^s}.
 \end{equation*}
+On the other hand, the variance term can be written as the sum of the variance of each empirical Fourier coefficient, and subsequently using the law of total variance, we have
+\begin{align}
+\label{pf:spectral_series_estimation_2}
+\Ebb\| \wt{f} - \Ebb[\wt{f}]\|_P^2 = \sum_{k = 1}^{K} \Var\Bigl[\dotp{{\bf Y}}{\psi_k}_n\Bigr] & = \sum_{k = 1}^{K} \Var\Bigl[\Ebb[\dotp{Y}{\psi_k}_n|{\bf X}\Bigr] + \Ebb\Bigl[\Var[\dotp{Y}{\psi_k}_n|{\bf X}\Bigr] \nonumber \\
+& = \sum_{k = 1}^{K} \Var\Bigl[\dotp{f_0}{\psi_k}_n\Bigr] + \frac{1}{n}\Ebb\Bigl[\|\psi_k\|_n^2\Bigr] \nonumber \\
+& \leq \frac{K}{n} + \sum_{k = 1}^{K}\Ebb\Bigl[\Bigl(f_0(X)\psi_k(X)\Bigr)^2\Bigr];
+\end{align}
+consequently,
+\begin{equation}
+\label{pf:spectral_series_estimation_1}
+\Ebb \|\wt{f} - f_0\|_P^2 \leq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\bigl[\lambda_{K + 1}(\Delta_P)\bigr]^s} + \frac{K}{n} + \frac{1}{n}\Ebb\Bigl[(f_0(X))^2 \cdot \sum_{k = 1}^{K} (\psi_k(X))^2\Bigr].
+\end{equation}
+The claim of the proposition then follows from two key facts. The first is a Weyl's Law scaling of the eigenvalues of $\Delta_P$, formally
+\begin{equation}
+\label{eqn:weyl}
+ck^{2/d} \leq \lambda_k(\Delta_P) \leq Ck^{2/d}\quad\textrm{for all $k \in \mathbb{N}\setminus \{1\}$},
+\end{equation}
+which is due to~\cite{dunlop2020}; the second is a local analog to Weyl's Law,
+\begin{equation}
+\label{eqn:local_weyl}
+\sup_{x \in \mc{X}}\biggl\{\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2\biggr\} \leq CK \quad\textrm{for all $K \in \mathbb{N}$},
+\end{equation}
+which follows straightforwardly from Theorem 17.5.3 of~\cite{hormander1973}.
+Plugging the upper bounds~\eqref{eqn:weyl} and~\eqref{eqn:local_weyl} back into~\eqref{pf:spectral_series_estimation_1} and applying the assumed upper bound $\Ebb[(f_0(X))^2] \leq 1$ yields
+\begin{equation*}
+\Ebb \|\wt{f} - f_0\|_P^2 \leq C\biggl(\frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{(k + 1)^s} + \frac{K}{n}\biggr),
+\end{equation*}
+and the choice $K = \floor{M^2n}^{d/(2s + d)} \vee 1$ then yields the claim.
+\qed.
+
+\paragraph{Testing.}
+In the goodness-of-fit testing problem, one asks for a test function---formally, a Borel measurable function $\phi$ that takes values in $\{0,1\}$--- which can distinguish between the hypotheses
+\begin{equation}
+\mathbf{H}_0: f_0 = f_0^{\star}, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{H}^{s}(\mc{X};M) \setminus \{f_0^{\star}\}.
+\end{equation} 
+To fix ideas, here and throughout we focus on the signal detection problem, which is the special case of $f_0^{\star} = 0$.\footnote{This is without loss of generality since all the test statistics we consider are easily modified to handle the case when $f_0^{\ast}$ is not $0$, by simply subtracting $f_0^{\ast}(X_i)$ from each observation $Y_i$, with no change in the analysis.}
+
+In this case, the classical spectral projection test $\wt{\varphi} = \1\{\wt{T} \geq \textcolor{red}{(?)}\}$ has bounded Type I error, $\Ebb_{0}[\wt{\varphi}] \leq a$.
+Proposition~\ref{prop:spectral_series_testing} gives an upper bound on the Type II error that holds uniformly over all $f_0 \in \mc{H}^s(\mc{X};M)$ which have a sufficiently large $L^2(P)$ norm.
+\begin{proposition}
+	\label{prop:spectral_series_testing}
+	Suppose data is observed according to Model~\ref{def:model_flat_euclidean}, and additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$ and $f_0 \in \mc{H}^{s}(\mc{X};M)$.
+Then there exists a constant $C$ which does not depend on $f_0,M$ or $n$ such that the following statement holds: if the spectral projection test $\wt{\varphi}$ is computed with parameter $K = \floor{M^2n}^{2d/(4s + d)}$, and if
+	\begin{equation}
+	\label{eqn:spectral_series_testing}
+	\|f_0\|_P^2 \geq \textcolor{red}{(?)}
+	\end{equation}
+	then the Type I error is bounded, $\Ebb_{f_0}[1 - \phi] \leq b$.
 \end{proposition}
-The proof of Proposition~\ref{prop:spectral_series_estimation}, along with proofs of all of our results, can be found in the appendix.
-It is worth mentioning some aspects of the analysis here, because it sets the stage for the strategy we will use to analyze Laplacian eigenmaps.
+Again, when the Sobolev ball radius $M \asymp 1$, the upper bound in~\eqref{eqn:spectral_series_testing} becomes $n^{-4s/(4s + d)}$, matching the usual minimax critical radius over Sobolev spaces (for more details on goodness-of-fit testing see~\citep{ingster2009,ingster2012}).
+% (AG 8/19/21): Same comment as above.
 
-In particular, three essential facts are needed to establish Proposition~\ref{prop:spectral_series_estimation}. 
-\begin{enumerate}
-	\item The continuous embedding of $H_0^s(\mc{X})$ into $\mc{H}^s(\mc{X})$---recall the latter is defined in~\eqref{eqn:spectral_sobolev_space}---which is a consequence of the zero-trace condition, the conditions on $\partial \mc{X}$ and $p$, and~\eqref{eqn:spectral_sobolev_space_equivalence}.
-	\item Weyl's Law, which gives the asymptotic scaling of eigenvalues $\lambda_{k}(\Delta_P) \asymp k^{2/d}$, and allows us to properly control the bias induced by spectral projection.
-	\item A local version of Weyl's law, which gives an estimate on $\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2$, and allows us to appropriately control the difference $\dotp{f_0}{\psi_k}_n - \dotp{f_0}{\psi_k}_P$  between the empirical and population Fourier coefficients.
-\end{enumerate}
-With these facts in hand the proof of Proposition~\ref{prop:spectral_series_estimation} follows from calculations standard to analysis of the Gaussian sequence model.
-Our analysis of Laplacian eigenmaps will depend on analogues of the first and second of these facts, with the space $\mc{H}^s(\mc{X})$ and eigenvalues $\lambda_k(\Delta_P)$ replaced by alternatives suitably defined with respect to the neighborhood graph Laplacian $L_{n,\varepsilon}$.
-
-\paragraph{Goodness-of-fit testing rates over Sobolev balls.}
-In the goodness-of-fit testing problem, we ask for a test function---formally, a Borel measurable function $\phi$ that takes values in $\{0,1\}$--- which can distinguish between the hypotheses
-\begin{equation}
-\mathbf{H}_0: f_0 = f_0^{\ast}, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{f_0^{\ast}\}.
-\end{equation} 
-Typically, the null hypothesis $f_0 = f_0^{\ast} \in \mc{F}$ reflects the absence of interesting structure, and $\mc{F} \setminus  \{f_0^{\ast}\}$ is a set of smooth departures from this null.
-To fix ideas, as in \citet{ingster2009} we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\ast} = 0$ and $\mc{F} = \mc{H}^s(\mc{X};M)$ is a Sobolev ball.
-This is without loss of generality since our test statistic and its analysis are easily modified to handle the case when $f_0^{\ast}$ is not $0$, by simply subtracting $f_0^{\ast}(X_i)$ from each observation $Y_i$.
-
-The Type I error of a test $\phi$ is $\mathbb{E}_0[\phi]$, and if $\mathbb{E}_0[\phi] \leq a$ for a given $a \in (0,1)$ we refer to $\phi$ as a level-$a$ test\footnote{We reserve the more common symbol $\alpha$ for multi-indices, so as to avoid confusion}.
-The worst-case risk of $\phi$ over $\mc{F}$ is
-\begin{equation*}
-R_n(\phi, \mc{F}, \epsilon) := \sup\Bigl\{\mathbb{E}_{f_0}[1 - \phi]: f_0 \in \mc{F}, \|f_0\|_{P} > \epsilon\Bigr\},
-\end{equation*}
-and for a given constant $b \in (0,1)$, the minimax critical radius $\epsilon_n(\mc{F})$ is the smallest value of $\epsilon$ such that some level-$a$ test has worst-case risk of at most $b$. Formally,
-\begin{equation*}
-\epsilon_n(\mc{F}) := \inf\Bigl\{\epsilon > 0: \inf_{\phi} R_n(\phi,\mc{F},\epsilon) \leq b \Bigr\},
-\end{equation*} 
-where in the above the infimum is over all level-$a$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Clearly, the minimax critical radius $\epsilon_n(\mc{F})$ depends on $a$ and $b$.
-However, we adopt the typical convention of treating $\alpha,b \in (0,1)$ and  as small but fixed positive constants; hence they will not affect the testing error rates, and we suppress them notationally.} We will refer to the rate at which the squared critical radius $\epsilon_n^2(\mc{F})$ tends to $0$ as the minimax testing rate.
-
-Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing rate is much smaller than the minimax estimation rate.
-\citet{ingster2009} give the minimax testing rate in the special case where $\mc{X} = [0,1]^d$ and $P$ is uniform, and we restate their result in the terms and notation of our paper.
-\begin{theorem}[Theorem 1 of \citet{ingster2009}]
-	\label{thm:spectral_series_testing}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $\mc{X} = [0,1]^d$ and $P$ is the uniform distribution on $[0,1]^d$.
-Then there exist constants $c,C,n_0$ which do not depend on $f_0$ or $n$ such that for all $n \geq n_0$,
-	\begin{equation}
-	\label{eqn:sobolev_space_testing_critical_radius}
-	c n^{-4s/(4s + d)} \leq \epsilon_n^2\bigl(\mc{H}^s(\mc{X};1)\bigr) \leq C n^{-4s/(4s + d)}~~\textrm{for $1 \leq d < 4s$.}
-	\end{equation}
-\end{theorem}
-The analysis used by \citet{ingster2009} to show the upper bound in~\eqref{eqn:sobolev_space_testing_critical_radius} relies on a similar trio of facts as used in the proof of Proposition~\ref{prop:spectral_series_estimation}.
-It is otherwise reminiscent of calculations made in the Gaussian sequence model, which can be found in~\cite{ingster2012}.
-This analysis can be straightforwardly adapted to handle design distributions $P$ that satisfy the conditions of Model~\ref{def:model_flat_euclidean}, or to handle the case where $M$ is not $1$.
-Finally, the test \citet{ingster2009} use to certify the upper bound in~\eqref{eqn:sobolev_space_testing_critical_radius} is implicitly a spectral projection method: the test statistic is the $L^2(P_n)$ norm of a particular spectral projection estimator.\footnote{\cite{ingster2009} project the responses onto a the span of trigonometric basis functions.
-Since $P$ is uniform on the unit cube, such functions are eigenfunctions of $\Delta_P$ when the right boundary conditions are imposed.}
-
-A major difference between testing and estimation over Sobolev spaces is the requirement that $4s > d$.
-When $4s \leq d$, the functions in $H^s(\Xset)$ are very irregular.
-Crucially, $H^s(\Xset)$ no longer continuously embeds into $\Leb^4(\Xset)$ when $4s \leq d$, and test statistics using the $L^2(P_n)$ norm are no longer guaranteed to have finite variance.\footnote{Note that this will not affect the analysis for estimation, because for estimation we only need to control the first two moments of $f_0$.} In fact, we have not seen any analysis which describes the minimax rate for nonparametric regression testing over Sobolev spaces in the $4s \leq d$ regime.
-However, if one explicitly assumes that $f_0 \in \Leb^4(\mc{X})$, then the critical radius is characterized by the dimension-free rate $\epsilon_n^2(\Leb^4(\mc{X};M)) \asymp Mn^{-1/2}$.\footnote{As a sanity check, note that when $1 \leq d < 4s$, $n^{-1/2}$ converges to $0$ at a slower rate than $n^{-4s/(4s + d)}$, which is the minimax rate for $M = 1$ given in~\eqref{eqn:sobolev_space_testing_critical_radius}.} As we discuss after our first main theorem regarding testing with Laplacian eigenmaps (Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}), this rate is achievable by a test based on $\wh{T}$.
-
-\paragraph{Manifold setup.}
-Under Model~\ref{def:model_manifold}, both~\eqref{eqn:sobolev_space_minimax_estimation_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced everywhere by the intrinsic dimension $m$.
-The estimator $\wt{f}$ and a test using the statistic $\wt{T}$---with $\psi_1,\psi_2,\ldots$ now the eigenfunctions of the manifold weighted Laplace-Beltrami operator $\Delta_P$---achieve the optimal estimation and testing rates.
-This is because each of the three facts mentioned after Proposition~\ref{prop:spectral_series_estimation} have analogues when the domain is a smooth manifold (See~\citet{hendriks1990}, who analyzes a spectral projection density estimator, for details.)
+The main takeaway from Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} is that spectral projection methods achieve optimal rates over Sobolev classes, when $p$ satisfies only a nonparametric notion of smoothness but is known.\footnote{The assumption $p \in C^{\infty}(\mc{X})$ could likely be weakened.
+Alternatively, one could likely derive upper bounds assuming Model~\ref{def:model_manifold} which depend only on the intrinsic dimension $m$.
+Since these would not substantially add to the main points of Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, we do not pursue the details further.} As we will shortly see, Laplacian eigenmaps achieves similar rates of convergence when $p$ is unknown, and in this sense both learns and leverages the design distribution in a way that more classical spectral projection methods cannot.
+It is worth pointing out that other methods besides Laplacian eigenmaps are statistically optimal for regression over Sobolev spaces even when the design distribution is unknown.
+We comment more on some of these in Section~\ref{sec:discussion}, after we have derived our major results regarding Laplacian Eigenmaps.
 

### 2021-08-26 manuscripts/laplacian_eigenmaps/sections/introduction.tex
Reworked proof of spectral projection estimator, moved into main text.

@@
 In contrast, in this paper we develop theory showing that graph spectral methods can effectively \emph{leverage} structure to perform downstream (supervised) learning tasks.
-Our main goal is to show that spectral projection procedures--which use eigenvectors of a neighborhood graph Laplacian as features in linear regression---are statistically optimal methods for nonparametric estimation and goodness-of-fit testing.
+Our main goal is to show that spectral projection procedures--which use eigenvectors of a neighborhood graph Laplacian as features in a linear regression problem---are statistically optimal methods for two problems in nonparametric regression: estimation and goodness-of-fit testing.
 
 The regression methods we consider are based on \emph{Laplacian eigenmaps}, first introduced by~\cite{belkin03a} \footnote{Technically speaking, as used by~\cite{belkin03a} ``Laplacian eigenmaps'' refers to a spectral embedding procedure.
-What we call Laplacian eigenmaps is a two-step regression method: the first step is this spectral embedding, and the second step computes the estimate $\wh{f}$ by running ordinary least squares in embedding space.
+What we call Laplacian eigenmaps is a two-step regression method: the first step is this spectral embedding, and the second step computes the estimator $\wh{f}$ by running ordinary least squares in embedding space.
 Since in this work we only ever consider the embedding within the context of this two-step regression method, we take the liberty of referring to the entire procedure as Laplacian eigenmaps.}.
-Given samples $(X_1,Y_1),\ldots,(X_n,Y_n)$, Laplacian eigenmaps first constructs a neighborhood graph---with vertices at the design points $X_1,\ldots,X_n$ and weighted edges between pairs of sufficiently close points $X_i$ and $X_j$---and computes the corresponding (unweighted) graph Laplacian.
-The Laplacian eigenmaps estimate $\wh{f}$ is found by projecting the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto a subspace spanned by a subset of the graph Laplacian eigenvectors, and the empirical norm $\|\wh{f}\|_n^2$ can be used as a statistic to test whether any signal is present: that is, whether the regression function $f_0(x) = \mathbb{E}[Y|X = x]$ is anywhere non-zero.
+Given samples $(X_1,Y_1),\ldots,(X_n,Y_n)$, Laplacian eigenmaps first forms a neighborhood graph, with vertices at the design points $X_1,\ldots,X_n$ and weighted edges between pairs of sufficiently close points $X_i$ and $X_j$.
+The Laplacian eigenmaps estimator $\wh{f}$ is then computed using the spectral decomposition of an (unweighted) Laplacian defined over this neighborhood graph, by projecting the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto a subspace spanned by a certain number of ``low-frequency'' graph Laplacian eigenvectors; more precisely, the eigenvectors corresponding to the smallest eigenvalues of the Laplacian.
+The Laplacian eigenmaps test statistic $\wh{T} = \|\wh{f}\|_n^2$ can in turn be used as a statistic to detect whether any signal is present, that is, whether the regression function $f_0(x) = \mathbb{E}[Y|X = x]$ is anywhere non-zero.
 (The precise definition of these methods is given in Section~\ref{subsec:laplacian_eigenmaps}.) We study these methods through the lens of a minimax analysis.
-Our major contribution is to provide upper bounds on the error of the Laplacian eigenmaps estimator and test which imply that they are minimax optimal over natural smoothness classes.
+Our major contribution is to provide upper bounds on the error of the Laplacian eigenmaps estimator and test which imply that they achieve the minimax optimal rate of convergence over natural smoothness classes.
 
@@
 The minimax estimation and testing rates for Sobolev spaces are generally speaking well-known---when using mean-squared error as the loss, the estimation rate is $n^{-2s/(2s + d)}$, and the testing rate is $n^{-4s/(4s + d)}$---and are achieved by various estimators.
-However, since Sobolev spaces admit a spectral characterization in terms of $\psi_1,\psi_2,\ldots$, classical spectral projection methods\footnote{We use ``classical'' to refer to spectral projection methods using eigenfunctions of a (fixed) Laplacian or Laplace-Beltrami operator, as distinct from Laplacian eigenmaps which uses eigenvectors of a (data-dependent) graph Laplacian matrix.} using these eigenfunctions are a particularly natural choice for regression over Sobolev spaces, and are in a sense canonical estimators for these function classes.
+However, since Sobolev spaces admit a spectral characterization in terms of $\psi_1,\psi_2,\ldots$, classical spectral projection methods\footnote{We use ``classical'' to refer to spectral projection methods using eigenfunctions of a (fixed) Laplacian or Laplace-Beltrami operator, as distinct from Laplacian eigenmaps which uses eigenvectors of a (data-dependent) graph Laplacian matrix.} using these eigenfunctions as features are a particularly natural choice for regression over Sobolev spaces, and are in a sense canonical estimators for these function classes.
 % (AG 8/13/21): Last clause makes a strong statement that I personally believe, but may irritate certain readers. Keep or remove?
@@
 In Section~\ref{sec:setup_main_results}, we give our formal modeling assumptions, and precisely define the Laplacian Eigenmaps methods we study.
-Propositions~\ref{prop:spectral_series_estimation} and ~\ref{prop:spectral_series_testing}, in Section~\ref{subsec:spectral_projection}, show that under rather general conditions on the design distribution, classical spectral projection methods achieve minimax rates of convergence over Sobolev classes.
-Then in Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps}, \ref{sec:out_of_sample} and and~\ref{sec:manifold_adaptivity}, we give our main upper bounds on the error of Laplacian eigenmaps (summarized above); these upper bounds certify that under similarly general conditions, Laplacian Eigenmaps is also minimax rate-optimal.
+Propositions~\ref{prop:spectral_series_estimation} and ~\ref{prop:spectral_series_testing}, in Section~\ref{subsec:spectral_projection}, show that under rather general (nonparametric) conditions on the design distribution, classical spectral projection methods achieve minimax rates of convergence over Sobolev classes.
+Then in Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps}, \ref{sec:out_of_sample} and and~\ref{sec:manifold_adaptivity}, we give our main upper bounds on the error of Laplacian eigenmaps (summarized above); these upper bounds hold under similarly general conditions, and imply that Laplacian Eigenmaps estimators and tests are also minimax rate-optimal.
 In Section~\ref{sec:experiments} we examine the empirical behavior of Laplacian eigenmaps, and show that even at moderate sample sizes Laplacian eigenmaps is competitive with classical spectral projection.

### 2021-08-26 manuscripts/laplacian_eigenmaps/sections/preliminaries.tex
Reworked proof of spectral projection estimator, moved into main text.

@@
 \end{equation}
-with regression function $f_0: \mc{X} \to \Reals$, and $w_i \sim N(0,1)$ independent Gaussian noise. 
+with regression function $f_0: \mc{X} \to \Reals$, and $w_i \sim N(0,1)$ independent Gaussian noise.)
 % (AG 8/4/21): Removed the comment regarding noise with variance = \sigma^2.
@@
 % (AG 8/13/21): I got rid of the sentence ``The data are sampled according to (1).'' It seemed repetitive. 
-At various points we will also assume that the density $p \in C^k(\mc{X})$ for some $k \in \mathbb{N}$; when we make such assumptions we will state them explicitly.
+At various points we will also assume that the density $p \in C^k(\mc{X})$; when we make such assumptions we will state them explicitly.
 
@@
 Boundary conditions plays an important role in the analysis of spectral methods, as we explain further in Section~\ref{subsec:spectral_projection}.
-For now, we limit ourselves to pointing out that for functions $f \in C^\infty(\mc{X})$, the zero-trace condition implies that $\partial^{k}f/\partial{\bf n}^k(x) = 0$ for each $k = 0,\ldots,s - 1$, and for all $x \in \partial\mc{X}$.
+For now, we limit ourselves to pointing out that for functions $f \in C^\infty(\mc{X})$, the zero-trace condition can be stated more concretely, as implying that $\partial^{k}f/\partial{\bf n}^k(x) = 0$ for each $k = 0,\ldots,s - 1$, and for all $x \in \partial\mc{X}$.
 (Here $\partial/(\partial {\bf n})$ is the partial derivative operator in the direction of the normal vector $\mathbf{n}$.)
@@
 
-The Laplacian eigenmaps estimator $\wh{f}$ simply projects the response vector ${\bf Y}$ onto the first $K$ eigenvectors of $L_{n,\varepsilon}$: letting $V_K \in \Reals^{n \times K}$ be the matrix with columns $v_1,\ldots,v_K$, we have that
+The Laplacian eigenmaps estimator $\wh{f}$ simply projects the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto the first $K$ eigenvectors of $L_{n,\varepsilon}$: letting $V_K \in \Reals^{n \times K}$ be the matrix with columns $V_{K,k} = v_k$, we have that
 \begin{equation}
@@
 
-\subsection{Classical Spectral Projection over Sobolev spaces}
+\subsection{Classical Spectral Projection}
 \label{subsec:spectral_projection}
-Laplacian Eigenmaps serves as a data-dependent alternative to more classical spectral projection methods constructed out of the eigenfunctions of a Laplace-Beltrami operator (or, in the special case where $\mc{X} = [0,1]^d$, out of a Fourier basis).
-We focus on the following density weighted Laplace-Beltrami operator $\Delta_P:C^2(\mc{X}) \to L^2(\mc{X})$,
+Laplacian Eigenmaps serves as a data-dependent alternative to more classical spectral projection methods.
+Classically, spectral projection methods operate by projecting the response vector ${\bf Y}$ onto a subspace spanned by eigenfunctions of some differential operator; for instance, the Laplacian operator $\Delta = \sum_{i = 1}^{d} \partial^2f/\partial x_i^2$, which corresponds to projection onto a Fourier basis.
+We will focus on a density-dependent analog to the Laplacian,  $\Delta_P:C^2(\mc{X}) \to L^2(\mc{X})$,
 \begin{equation}
@@
 \end{equation}
-with eigenvalue/eigenfunction pairs $(\lambda_1(\Delta_P),\psi_1),(\lambda_2(\Delta_P),\psi_2),\ldots$ defined as solutions to
+which has enumerable eigenvalue/eigenfunction pairs $(\lambda_k(\Delta_P),\psi_k),(\lambda_2(\Delta_P),\psi_2),\ldots$ defined according to
 \begin{equation}
@@
 \end{equation}
-and sorted as usual in ascending order of eigenvalue.\footnote{For formal justification of why~\eqref{eqn:laplace_beltrami_eigenproblem} has discrete spectrum under either Model~\ref{def:model_flat_euclidean} or~\ref{def:model_manifold}, see~\cite{garciatrillos18,trillos2019}.} In this case, classical spectral projection refers to the estimator and test statistic
+and sorted as usual in ascending order of eigenvalue.\footnote{For formal justification of why~\eqref{eqn:laplace_beltrami_eigenproblem} has discrete spectrum under either Model~\ref{def:model_flat_euclidean} or~\ref{def:model_manifold}, see~\cite{garciatrillos18,trillos2019}.} The classical spectral projection estimator and test statistic are then simply
 \begin{equation}
@@
 \end{equation}
-The statistical behavior of these more classical methods is relevant because the spectrum of $\Delta_P$ serves as the population-level limit of graph Laplacian spectra: for any fixed $k \in \mathrm{N}$, as $n \to \infty$
+We focus on the particular differential operator $\Delta_P$ because the spectrum of $\Delta_P$ is the population-level limit of graph Laplacian spectra: for any fixed $k \in \mathbb{N}$, as the number of samples $n \to \infty$ and the graph radius $\varepsilon \to 0$,
 \begin{equation}
-\lambda_k \to \lambda_k(\Delta_P), \quad\textrm{and}\quad \Bigl\|\frac{1}{n}v_k - \psi_k\Bigr\|_n^2 \to 0,
+\lambda_k \to \lambda_k(\Delta_P), \quad\textrm{and}\quad \Bigl\|\frac{1}{n}v_k - \psi_k\Bigr\|_n^2 \to 0.
 \end{equation}
-in either the flat Euclidean (Model~\ref{def:model_flat_euclidean}) or manifold (Model~\ref{def:model_manifold}) setups~\citep{garciatrillos18,trillos2019}.
-It follows that for any fixed $K \in \mathbb{N}$, the Laplacian eigenmaps estimator converges to the $\wh{f} \to \wt{f}$, and likewise $\wh{T} \to \wt{T}$, justifying why we view Laplacian Eigenmaps as a data-dependent alternative to classical spectral projection.
+This holds in either the flat Euclidean (Model~\ref{def:model_flat_euclidean}) or manifold (Model~\ref{def:model_manifold}) setups~\citep{garciatrillos18,trillos2019}.
+It follows that for any fixed number of eigenvectors $K \in \mathbb{N}$, both $\wh{f} \to \wt{f}$ and $\wh{T} \to \wt{T}$, and in this way Laplacian Eigenmaps can be formally tied to classical spectral projection methods for regression.
 
-On the other hand, the eigenvectors and eigenfunctions of a (possibly density-weighted) Laplace-Beltrami operator such as $\Delta_P$ can also be used to give a spectral definition of Sobolev spaces.
-For instance, consider the ellipsoid
+On the other hand, the eigenvectors and eigenfunctions of $\Delta_P$ can also be used to define Sobolev spaces in an alternative manner. Consider the ellipsoid
 \begin{equation}
@@
 \end{equation}
-with corresponding norm $\|\sum_{k = 1}^{\infty} a_k \psi_k\|_{\mc{H}^s(\mc{X})}^2 = \sum_{k = 1}^{\infty} a_k^2 \lambda_{k}(\Delta_P)$.
+equipped with the norm $\|\sum_{k = 1}^{\infty} a_k \psi_k\|_{\mc{H}^s(\mc{X})}^2 = \sum_{k = 1}^{\infty} a_k^2 \lambda_{k}(\Delta_P)$.
 Under appropriate regularity conditions $\mc{H}^s(\mc{X})$ consists of functions $f \in H^s(\mc{X})$ which also satisfy some additional boundary conditions.
-For instance, assuming Model~\ref{def:model_flat_euclidean}, $p \in C^{\infty}(\mc{X})$ and $\partial \mc{X} \in C^{1,1}$, \citet{dunlop2020} show that the ellipsoid $\mc{H}^{2s}(\mc{X})$ satisfies
+For instance, assuming Model~\ref{def:model_flat_euclidean}, $p \in C^{\infty}(\mc{X})$ and $\partial \mc{X} \in C^{1,1}$, \citet{dunlop2020} show that for any $s \geq 1$, the ellipsoid $\mc{H}^{2s}(\mc{X})$ satisfies
 \begin{equation}
@@
 \end{equation}
-for any $s > 0$, and likewise $\mc{H}^{2s + 1}(\mc{X}) = \mc{H}^{2s}(\mc{X}) \cap H^{2s + 1}(\mc{X})$ for any $s \geq 0$; additionally, the norms $\|\cdot\|_{\mc{H}^s(\mc{X})}$ and $\|\cdot\|_{H^s(\mc{X})}$ are equivalent.
+and likewise $\mc{H}^{2s + 1}(\mc{X}) = \mc{H}^{2s}(\mc{X}) \cap H^{2s + 1}(\mc{X})$ for any $s \geq 0$; additionally, the norms $\|\cdot\|_{\mc{H}^s(\mc{X})}$ and $\|\cdot\|_{H^s(\mc{X})}$ are equivalent.
 
-The spectral formulation of Sobolev spaces suggest that classical spectral projection methods---and in turn Laplacian Eigenmaps---are natural choices for regression over such function classes.
-Further, given the similarities between Laplacian Eigenmaps and classical spectral projection methods, the statistical properties of the latter---which are better understood and easier to derive---should intuitively shed some light on the properties of the former.
-To this end, we now give a pair of results showing that classical spectral methods are statistically optimal for nonparametric regression over Sobolev classes with random design, \emph{so long as the design distribution is known}.
-These results are in line with previously known upper bounds on the estimation and testing error of classical spectral methods~\citep{tsybakov08,ingster2009}, but hold under less stringent conditions on the design distribution.
-This latter point is important, since our interest in Laplacian Eigenmaps stems in part from its adaptivity to general and unknown designs.
+This spectral formulation suggests that classical spectral projection methods---and in turn Laplacian Eigenmaps---are natural choices for regression over Sobolev spaces.
+Furthermore, since classical spectral projection methods serve, in a sense, as the infinite-data limit of Laplacian eigenmaps, the statistical properties of the former---which are generally speaking better understood and easier to derive---should intuitively shed some light on the properties of the latter.
+To this end, we now give a pair of results, Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, which establish that classical spectral projection methods are statistically optimal for nonparametric estimation and testing over Sobolev ellipsoids.
+These results are generalizations of previously known upper bounds on the estimation and testing error of classical spectral methods~\citep{tsybakov08,ingster2009}, in that they hold under less stringent (nonparametric) conditions on the design distribution $P$.
+This latter point is important, since as we have argued one attractive quality of Laplacian Eigenmaps is its adaptivity to general design distributions.
 
@@
 	\label{prop:spectral_series_estimation}
-	Suppose data is observed according to Model~\ref{def:model_flat_euclidean}, and additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$, $f_0 \in \mc{H}^{s}(\mc{X};M)$ and $\|f_0\|_P^2 \leq 1$.
+	Suppose data is observed according to Model~\ref{def:model_flat_euclidean}, and that the density $p$ is known.
+Suppose additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$, $f_0 \in \mc{H}^{s}(\mc{X};M)$ and $\|f_0\|_P^2 \leq 1$.
 Then there exists a constant $C$ which does not depend on $f_0,M$ or $n$ such that the following statement holds: if the spectral projection estimator $\wt{f}$ is computed with parameter $K = \floor{M^2n}^{d/(2s + d)} \vee 1$, then
@@
 
-We now give the proof of Proposition~\ref{prop:spectral_series_estimation}; the structure of the analysis, which is fairly classical and straightforward, is similar to the high-level strategy we use to analyze Laplacian Eigenmaps (see Section~\ref{subsec:analysis}).
+We now give the proof of Proposition~\ref{prop:spectral_series_estimation}.
+The structure of the analysis, which is fairly classical and straightforward, can be usefully compared and contrasted with our analysis of Laplacian Eigenmaps (see Section~\ref{subsec:analysis}).
 
@@
 \end{equation}
-Since the eigenfunctions $\{\psi_k\}$ form an orthonormal basis of $L^2(P)$, and $f_0 \in H_0^{s}(\mc{X}) \subseteq L^2(P)$, we can write the squared bias in terms of squared Fourier coefficients of $f_0$, leading to the following upper bound,
+Since the eigenfunctions $\{\psi_k\}$ form an orthonormal basis of $L^2(P)$, and $f_0 \in \mc{H}^{s}(\mc{X}) \subseteq L^2(P)$, we can write the squared bias in terms of squared Fourier coefficients of $f_0$, leading to the following upper bound,
 \begin{equation*}
-\|f_0 - \Ebb{\wt{f}}\|_P^2 = \sum_{k = K + 1}^{\infty}  \dotp{f_0}{\psi_k}^2 \leq  \frac{1}{\lambda_{K + 1}(\Delta_P)^s} \sum_{k = K + 1}^{\infty} \lambda_{k + 1}(\Delta_P)^s \dotp{f_0}{\psi_k}^2 \leq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}}{\bigl[\lambda_{K + 1}(\Delta_P)\bigr]^s}.
+\|f_0 - \Ebb[\wt{f}]\|_P^2 = \sum_{k = K + 1}^{\infty}  \dotp{f_0}{\psi_k}_P^2 \leq  \frac{1}{\{\lambda_{K + 1}(\Delta_P)\}^s} \sum_{k = K + 1}^{\infty} \{\lambda_{k + 1}(\Delta_P)\}^s \dotp{f_0}{\psi_k}_P^2 \leq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}}{\{\lambda_{K + 1}(\Delta_P)\}^s}.
 \end{equation*}
-On the other hand, the variance term can be written as the sum of the variance of each empirical Fourier coefficient, and subsequently using the law of total variance, we have
+On the other hand, the variance term can be written as the sum of the variance of each empirical Fourier coefficient, and subsequently by the law of total variance we derive that
 \begin{align}
@@
 & = \sum_{k = 1}^{K} \Var\Bigl[\dotp{f_0}{\psi_k}_n\Bigr] + \frac{1}{n}\Ebb\Bigl[\|\psi_k\|_n^2\Bigr] \nonumber \\
-& \leq \frac{K}{n} + \sum_{k = 1}^{K}\Ebb\Bigl[\Bigl(f_0(X)\psi_k(X)\Bigr)^2\Bigr];
+& \leq \frac{K}{n} + \frac{1}{n}\sum_{k = 1}^{K}\Ebb\Bigl[\Bigl(f_0(X)\psi_k(X)\Bigr)^2\Bigr].
 \end{align}
-consequently,
+Consequently,
 \begin{equation}
@@
 \end{equation}
-The claim of the proposition then follows from two key facts. The first is a Weyl's Law scaling of the eigenvalues of $\Delta_P$, formally
+The claim of the proposition then follows from two (variants of) classical results in spectral geometry.
+The first is a Weyl's Law asymptotic scaling of the eigenvalues of $\Delta_P$ due to~\cite{dunlop2020}; formally, there exist constants $c$ and $C$ (which will depend on $P$ and $d$) such that
 \begin{equation}
 \label{eqn:weyl}
-ck^{2/d} \leq \lambda_k(\Delta_P) \leq Ck^{2/d}\quad\textrm{for all $k \in \mathbb{N}\setminus \{1\}$},
+ck^{2/d} \leq \lambda_k(\Delta_P) \leq Ck^{2/d}\quad\textrm{for all $k \in \mathbb{N}$, $k \geq 2$}.
 \end{equation}
-which is due to~\cite{dunlop2020}; the second is a local analog to Weyl's Law,
+The second is a local analog to Weyl's Law,
 \begin{equation}
@@
 which follows straightforwardly from Theorem 17.5.3 of~\cite{hormander1973}.
-Plugging the upper bounds~\eqref{eqn:weyl} and~\eqref{eqn:local_weyl} back into~\eqref{pf:spectral_series_estimation_1} and applying the assumed upper bound $\Ebb[(f_0(X))^2] \leq 1$ yields
+Plugging the upper bounds~\eqref{eqn:weyl} and~\eqref{eqn:local_weyl} back into~\eqref{pf:spectral_series_estimation_1} and applying the assumed upper bound $\Ebb[(f_0(X))^2] \leq 1$ shows that
 \begin{equation*}
@@
 \end{equation*}
-and the choice $K = \floor{M^2n}^{d/(2s + d)} \vee 1$ then yields the claim.
+and setting $K = \floor{M^2n}^{d/(2s + d)} \vee 1$, which balances squared bias and variance, then yields the claim.
 \qed.
@@
 
-In this case, the classical spectral projection test $\wt{\varphi} = \1\{\wt{T} \geq \textcolor{red}{(?)}\}$ has bounded Type I error, $\Ebb_{0}[\wt{\varphi}] \leq a$.
+In this case, the classical spectral projection test $\wt{\varphi} = \1\{\wt{T} \geq K/N + \sqrt{K/a}/n\}$ has bounded Type I error, $\Ebb_{0}[\wt{\varphi}] \leq a$.
 Proposition~\ref{prop:spectral_series_testing} gives an upper bound on the Type II error that holds uniformly over all $f_0 \in \mc{H}^s(\mc{X};M)$ which have a sufficiently large $L^2(P)$ norm.
@@
 	\label{prop:spectral_series_testing}
-	Suppose data is observed according to Model~\ref{def:model_flat_euclidean}, and additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$ and $f_0 \in \mc{H}^{s}(\mc{X};M)$.
+	Suppose data is observed according to Model~\ref{def:model_flat_euclidean}, and that the density $p$ is known.
+Suppose additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$ and $f_0 \in \mc{H}^{s}(\mc{X};M)$.
 Then there exists a constant $C$ which does not depend on $f_0,M$ or $n$ such that the following statement holds: if the spectral projection test $\wt{\varphi}$ is computed with parameter $K = \floor{M^2n}^{2d/(4s + d)}$, and if
@@
 
-The main takeaway from Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} is that spectral projection methods achieve optimal rates over Sobolev classes, when $p$ satisfies only a nonparametric notion of smoothness but is known.\footnote{The assumption $p \in C^{\infty}(\mc{X})$ could likely be weakened.
+The main takeaway from Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} is that spectral projection methods achieve optimal rates over Sobolev classes, when the design density $p$ satisfies only a nonparametric notion of smoothness\footnote{The assumption $p \in C^{\infty}(\mc{X})$ could likely be weakened.
 Alternatively, one could likely derive upper bounds assuming Model~\ref{def:model_manifold} which depend only on the intrinsic dimension $m$.
-Since these would not substantially add to the main points of Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, we do not pursue the details further.} As we will shortly see, Laplacian eigenmaps achieves similar rates of convergence when $p$ is unknown, and in this sense both learns and leverages the design distribution in a way that more classical spectral projection methods cannot.
-It is worth pointing out that other methods besides Laplacian eigenmaps are statistically optimal for regression over Sobolev spaces even when the design distribution is unknown.
+Since these would not substantially add to the main points of Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, we do not pursue the details further.} but, crucially, is known a priori.
+As we will see, Laplacian eigenmaps achieves similar rates of convergence when $p$ is unknown, and in this sense both learns and leverages the design distribution in a way that more classical spectral projection methods cannot.
+It is worth pointing out that other methods besides Laplacian eigenmaps are statistically optimal for nonparametric regression even when $p$ is unknown.
 We comment more on some of these in Section~\ref{sec:discussion}, after we have derived our major results regarding Laplacian Eigenmaps.

### 2021-08-28 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Through upper bounds on Laplacian eigenmaps in flat Euclidean case.

@@
 
-As previously explained, Laplacian eigenmaps is a discrete and noisy approximation to a spectral projection method using the eigenfunctions of $\Delta_P$.
+% As previously explained, Laplacian eigenmaps is a discrete and noisy approximation to a spectral projection method using the eigenfunctions of $\Delta_P$.
 This is particularly useful when $P$ is unknown, or when the eigenfunctions of $\Delta_P$ cannot be explicitly computed.
@@
 In this section and the following one, we will see that this is indeed the case: the estimator $\wh{f}$, and a test using the statistic $\wh{T}$, achieve optimal estimation and goodness-of-fit testing rates over Sobolev classes.
-
-In this section we will cover the flat Euclidean case, where we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to Model~\ref{def:model_flat_euclidean}.
-We will divide our theorem statements based on whether we assume the regression function $f_0$ belongs to the first order Sobolev class ($s = 1$) or a higher-order Sobolev class ($s > 1$), since the details of the two settings are somewhat different.
+% (AG 8/26/21): This segment is now redundant.
+
+In this section we give upper bounds on the error of Laplacian Eigenmaps in the flat Euclidean case, where we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to Model~\ref{def:model_flat_euclidean}.
+We will divide our theorem statements based on the regression function $f_0$ belongs to the first order Sobolev class $H^1(\mc{X})$ or a higher-order Sobolev class ($H_0^{s}(\mc{X})$ for some $s > 1$), since the details of the two settings are somewhat different.
 
@@
 We begin by assuming $f_0 \in H^1(\mc{X}; M)$.
-We show that $\wh{f}$ and a test based on $\wh{T}$ are minimax optimal, for all values of $d$, and under no additional assumptions (beyond those of Model~\ref{def:model_flat_euclidean}) on the data generating process, i.e.
-on either $P$ or $f_0$.
-
-\paragraph{Estimation.} When the kernel $\eta$, graph radius $\varepsilon$, and number of eigenvectors $K$ are chosen appropriately, we show in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that the estimator $\wh{f}$ achieves the minimax rate over $H^1(\mc{X};M)$, when the error is measured in squared $L^2(P_n)$ norm.
+We show that $\wh{f}$ and a test based on $\wh{T}$ are minimax optimal, for all values of $d$ for which the minimax rates are known, and under no additional assumptions (beyond those of Model~\ref{def:model_flat_euclidean}) on the design distribution $P$.
+
+\paragraph{Estimation.} Laplacian eigenmaps depends on the kernel $\eta$ and two tuning parameters, the graph radius $\varepsilon$ and number of eigenvectors $K$. We now give some assumptions on each.
 \begin{enumerate}[label=(K\arabic*)]
@@
 \end{theorem}
-From~\eqref{eqn:laplacian_eigenmaps_estimation_fo} it follows immediately that with constant probability $\|\wh{f} - f_0\|_n^2 \lesssim M^2(M^2n)^{-2/(2 + d)}$ whenever $n^{-1/2} \lesssim M \lesssim n^{1/d}$, matching the standard minimax estimation rate over Sobolev classes.
+From~\eqref{eqn:laplacian_eigenmaps_estimation_fo} it follows immediately that when $M \asymp 1$, then with constant probability $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2/(2 + d)}$, matching the standard minimax estimation rate over Sobolev classes.
 
@@
 \begin{itemize}
-	\item When $M = o(n^{-1/2})$, then computing Laplacian eigenmaps with $K = 1$ achieves the parametric rate $\|\wh{f} - f_0\|_n^2 \lesssim n^{-1}$, and the zero-estimator $\wh{f} = 0$ achieves the better rate $\|\wh{f} - f_0\|_n^2 \lesssim M^2$.
+	\item \emph{Radius of the Sobolev ball.} If $M = M_n$ satisfies $n^{-1/2} \lesssim M_n \lesssim n^{1/d}$, then Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} implies that $\|\wh{f} - f_0\|_n^2 \lesssim M^2(M^2n)^{-2/(2 + d)}$ with constant probability.
+This matches the upper bound on $\Ebb\|\wt{f} - f_0\|_P^2$ given in~\eqref{eqn:spectral_series_estimation}, including in its dependence on $M$.
+It also matches the known minimax rate for some related problems, including estimation in the Gaussian sequence model~\citep{johnstone2011}, and estimation with a fixed design when the dimension $d = 1$~\citep{vandergeer2000}.
+
+	When $M = o(n^{-1/2})$ then computing Laplacian eigenmaps with $K = 1$ achieves the parametric rate $\|\wh{f} - f_0\|_n^2 \lesssim n^{-1}$, and the zero-estimator $\wh{f} = 0$ achieves the better rate $\|\wh{f} - f_0\|_n^2 \lesssim M^2$.
 However, we do not know what the minimax rate is in this regime.
-On the other hand, when $M = \omega(n^{1/d})$, then computing Laplacian eigenmaps with $K = n$ achieves the rate $\|\wh{f} - f_0\|_n^2 \lesssim 1$, which is better than the rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}.
-This is because we are evaluating error in $L^2(P_n)$ rather than $L^2(P)$.
+On the other hand, when $M = \omega(n^{1/d})$, then computing Laplacian eigenmaps with $K = n$ achieves the rate $\|\wh{f} - f_0\|_n^2 \lesssim 1$, which is better than the rate in~\eqref{eqn:spectral_series_estimation}.
+This is because we are evaluating error in-sample rather than out-of-sample.
 However, in truth these are edge cases, which do not fall neatly into the framework of nonparametric regression.
-	\item The assumptions placed on the kernel function $\eta$ are needed for technical reasons.
+
+	\item \emph{In-sample error.} As we've already noted, since the Laplacian eigenmaps estimator is defined only at the design points $\{X_1,\ldots,X_n\}$, we will use the empirical norm $\|\cdot\|_n^2$ as our estimation loss, and return to the question of out-of-sample estimation later in Section~\ref{sec:out_of_sample}.
+
+	There is one subtlety introduced by the use of in-sample mean squared error.
+Since elements $f \in H^s(\mc{X})$ are equivalence classes, defined only up to a set of measure zero, one cannot really speak of the pointwise evaluation $f_0(X_i)$, as we do by defining our target of estimation to be $(f_0(X_1),\ldots,f_0(X_n))$, until one selects a representative of each equivalence class $f$.
+Implicitly, we will always pick the \emph{precise representative} $f_0^{\ast} \in f_0$ (as defined in~\cite{evans15}), and the notation ``$f_0(X_i)$'' should always be interpreted as $f_0^{\ast}(X_i)$.
+To be clear, however, it does not really matter which representative we choose, since all versions agree except on a set of measure zero, and so any two $g_0,h_0 \in f_0$ satisfy $g_0(X_i) = h_0(X_i)$ for all $i = 1,\ldots,n$ almost surely.
+For this reason we can write $f_0(X_i)$ without fear of ambiguity or confusion.
+
+	\item \emph{Tuning parameters}.
+The assumptions placed on the kernel function $\eta$ are needed for technical reasons.
 They can likely be weakened, although we note that they are already fairly general.
-The lower bound on $\varepsilon$ imposed by~\eqref{eqn:radius_fo} is on the order of the connectivity threshold, the smallest length scale at which the resulting graph will still be connected with high probability.
-On the other hand, as we will see in Section~\ref{subsec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K$ is of at least the same order as the continuum eigenvalue $\lambda_K(\Delta_P)$.
-Finally, we choose $K = (M^2n)^{d/(2 + d)}$ (when possible) to optimally trade-off bias and variance.
-	\item We note that the ranges~\eqref{eqn:radius_fo} and~\eqref{eqn:eigenvector_estimation_fo} depends on the radius of the Sobolev ball $M$, which is usually unknown.
-In practice, one typically tunes hyper-parameters by sample-splitting or cross-validation.
-However, because the estimator $\wh{f}$ is defined only in-sample, we cannot use such methods to select the graph radius $\varepsilon$, or number of eigenvectors $K$.
+The lower bound on $\varepsilon$ imposed by~\eqref{eqn:radius_fo} is on the order of the connectivity threshold, the smallest radius for which the resulting graph will still be connected with high probability.
+On the other hand, as we will see in Section~\ref{subsec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K$ is of at least the same order as the continuum eigenvalue $\lambda_K(\Delta_P)$; this is essential in order to obtain a tight upper bound on the bias of $\wh{f}$.
+Finally, we set $K = \floor{(M^2n)^{d/(2 + d)}}$ (when possible) to optimally trade-off bias and variance.
+
+	In practice, one typically tunes hyper-parameters by cross-validation.
+However, because the estimator $\wh{f}$ is defined only in-sample, neither cross-validation nor any other sample-splitting method can be used to tune parameters for Laplacian eigenmaps.
 We return to this issue in Section~\ref{sec:out_of_sample}, when we propose an out-of-sample extension of $\wh{f}$.
 
-	\textcolor{red}{(TODO): Ryan, you asked about AIC, but I'm not clear about whether you want me to (a) make some remark about it, (b) actually go through the work of giving guarantees, or (c) were just curious.}
-
-	\item The upper bound given in equation~\eqref{eqn:laplacian_eigenmaps_estimation_fo} holds with probability $1 - \delta - Cn\exp(-cn\varepsilon^d)$.
-Under the stronger assumption that $f_0 \in C^1(\mc{X};M)$ we can replace the factor of $\delta$ by the sharper $\delta^2/n$, which is less than $\delta$ because $\delta \in (0,1)$.
-Then a routine calculation shows that the expected $L^2(P_n)$ loss is on the same order as~\eqref{eqn:laplacian_eigenmaps_estimation_fo}, matching the minimax rate.
+	% (AG 8/27/21): Ryan, you asked about AIC, but I'm not clear about whether you want me to (a) make some remark about it, (b) actually go through the work of giving guarantees, or (c) were just curious.
+	\item \emph{High-probability guarantees}.
+The upper bound given in~\eqref{eqn:laplacian_eigenmaps_estimation_fo} holds with ``constant probability'', meaning with probability $1 - \delta - o(1)$.
+Under the stronger assumption that $f_0$ is $M$-Lipschitz, we can establish the same guarantee~\eqref{eqn:laplacian_eigenmaps_estimation_fo} with probability $1 - \delta^2/n - Cn\exp(-cn\varepsilon^d) - \exp(-K)$; in other words, we can give a high probability guarantee (for details see~\citep{green2021}).
+In this case a routine calculation shows that $\Ebb[\|\wh{f} - f_0\|_n^2]$ will also be on the some order as~\eqref{eqn:laplacian_eigenmaps_estimation_fo}.
+We also suspect that high-probability guarantees will hold so long as $\|\nabla f\|_{L^q(\mc{X})}$ is bounded for some sufficiently large $q < \infty$, but it remains an open question whether such guarantees can be obtained in the Sobolev case ($q = 2$) which is the focus of this work.
 \end{itemize}
@@
 This choice of threshold $t_{a}$ guarantees that $\varphi$ is a level-$a$ test.
-As we show in Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}, when $d < 4$, $\varepsilon$ and $K$ are chosen appropriately, and the alternative $f_0$ is separated from $0$ by at least (a constant times) the minimax critical radius $M^2(M^2n)^{-4/(4 + d)}$, the test $\varphi$ has Type II error of at most $b$.
+As we show in Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}, when $d < 4$, $\varepsilon$ and $K$ are chosen appropriately, and the alternative $f_0$ has is sufficiently well-separated from $0$, the test $\varphi$ has Type II error of at most $b$.
 
@@
 Although~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} involves taking the maximum of several different terms, the important takeaway of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} is that if $n^{-1/2} \lesssim M \lesssim n^{(4 - d)/4d}$, then $\varphi$ has small worst-case risk as long as $f_0$ is separated from $0$ by at least $M^2(M^2n)^{-4/(4 + d)}$.
-Note that unlike in the estimation setting---where we measured loss in $L^2(P_n)$ error---the separation in~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} is measured in $L^2(P)$ norm.
-Thus~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} implies that $\varphi$ is a minimax rate-optimal test over $H^1(\mc{X};M)$, in the usual sense.
-
-As mentioned previously, when $d \geq 4$ the first order Sobolev space $H^1(\mc{X})$ does not continuously embed into $\Leb^4(\mc{X})$, and we do not know the optimal rates for regression testing over $H^1(\mc{X};M)$.
-On the other hand, if we explicitly assume $f_0 \in \Leb^4(\mc{X})$, then the Laplacian eigenmaps test with $K = n$, has small type II error whenever $\|f_0\|_P^2 \gtrsim Mn^{-1/2}$.
-Note that when $K = n$ the Laplacian eigenmaps test statistic is nothing but the squared $L^2(P_n)$ norm of ${\bf Y}$, $\wh{T} = \|{\bf Y}\|_n^2$.
-See \textcolor{blue}{(Green et al.
-2021)} for details.
+In the special case $M \asymp 1$, this matches the minimax squared critical radius $n^{-4/(4 + d)}$, implying that $\varphi$ is a minimax rate-optimal test over $H^1(\mc{X};M)$ when $d \in \{1,2,3\}$.
+As mentioned previously, when $d \geq 4$ the first order Sobolev space $H^1(\mc{X})$ does not continuously embed into $\Leb^4(\mc{X})$, and in this case the optimal rates for regression testing over Sobolev spaces are unknown.
 
@@
 \label{sec:higher_order_sobolev_classes}
-We now consider the situation where the regression function displays some higher-order regularity.
-For reasons already discussed, we also assume the zero-trace boundary condition, i.e.
-$f_0 \in H_0^s(\mc{X})$.
+We now consider the situation where the regression function displays some higher-order regularity, $f_0 \in H_0^s(\mc{X})$.
 We show that Laplacian eigenmaps methods continue to be optimal for all orders of $s$, as long as the design density is itself also sufficiently regular, $p \in C^{s - 1}(\mc{X})$.
@@
 \end{theorem}
-Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that in the flat Euclidean setting Laplacian eigenmaps is an in-sample minimax rate-optimal estimator over Sobolev classes, for all values of $s$ and $d$.
+Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that in the flat Euclidean setting Laplacian eigenmaps is a minimax rate-optimal estimator over Sobolev classes, for all values of $s$ and $d$.
 Some other remarks:
 \begin{itemize}
-	\item We do not require that the regularity of the Sobolev space satisfy $s > d/2$, a condition often seen in the literature.
+	\item \emph{Sub-critical Sobolev spaces}.
+We do not require that the regularity of the Sobolev space satisfy $s > d/2$, a condition often seen in the literature.
 In the sub-critical regime $s \leq d/2$, the Sobolev space $H^s(\mc{X})$ is quite irregular.
@@
 Likewise, certain estimators are ``off the table'', most notably RKHS-based methods such as thin-plate splines of degree $k \leq d/2$.
-Nevertheless, for random design regression with error measured in $\Leb^2(P)$-norm, the spectral projection estimator $\wt{f}$ defined in~\eqref{eqn:spectral_series_estimator} obtains the ``usual'' minimax rates $n^{-2s/(2s + d)}$ for all values of $s$ and $d$.
+Nevertheless, for random design regression with error measured in $\Leb^2(P)$-norm, the spectral projection estimator $\wt{f}$ obtains the standard minimax rates $n^{-2s/(2s + d)}$ for all values of $s$ and $d$.
 Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and \ref{thm:laplacian_eigenmaps_estimation_ho} show that the same is true with respect to Laplacian eigenmaps, with error measured in $\Leb^2(P_n)$-norm.
-	\item The requirement $p \in C^{s - 1}(\mc{X})$ is a strong condition, but is essential to showing that $\wh{f}$ enjoys faster rates of convergence when $s > 1$.
-We explain why in Section~\ref{subsec:analysis}, where we discuss our analysis.
+	\item \emph{Smoothness of design density}.
+As promised,  Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} shows that Laplacian eigenmaps achieves optimal rates of convergence so long as the unknown design density $p$ is sufficiently smooth, $p \in C^{s - 1}(\mc{X})$.
+The requirement $p \in C^{s - 1}(\mc{X})$ is essential to showing that $\wh{f}$ enjoys the faster minimax rates of convergence when $s > 1$,  as we discuss in Section~\ref{subsec:analysis}.
 \end{itemize}
@@
 \end{theorem}
-Similarly to the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that $\varphi$ is a minimax optimal test over $H_0^s(\mc{X})$ when $n^{-1/2} \lesssim M \lesssim n^{(4s - d)/4d}$.
-However, unlike the first-order case, when $4 < d < 4s$ the minimax testing rate over $H_0^s(\mc{X})$ is still on the order of $M^2(M^2n)^{-4s/(4s + d)}$.
-Unfortunately, we can no longer claim that $\varphi$ is an optimal test in this regime.
+Similarly to the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that when $M \asymp 1$, then $\varphi$ is a minimax rate-optimal test over $H_0^s(\mc{X})$.
+However, unlike the first-order case, when $4 < d < 4s$ the minimax testing rate over $H_0^s(\mc{X})$ is still on the order of $M^2(M^2n)^{-4s/(4s + d)}$; unfortunately, we can no longer claim that $\varphi$ is an optimal test in this regime.
 \begin{theorem}
@@
 \end{theorem}
-Note that as a consequence of Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, if we choose $K = n^{-2s/(2s + d)}$ then $\varphi$ must have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2s + d)}$.
-Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} shows that $\varphi$ can achieve better, but still not optimal, rates.
-As a technical matter, the problem is that when $d > 4$ there do not exist any choices of $\varepsilon$ and $K$ which satisfy both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho}, and as a result we cannot optimally balance (our upper bound on) testing bias and variance (defined momentarily in~\eqref{eqn:testing_biasvariance}).
-Although we suspect $\varphi$ is truly suboptimal when $d > 4$, technically speaking~\eqref{eqn:testing_biasvariance} gives only an upper bound on testing bias, and thus we cannot rule out that the test $\varphi$ is optimal for all $4 < d < 4s$.
+Focusing again on the special case where $M \asymp 1$, Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} says that $\varphi$ has have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2(s - 1) + d)}$ and $4 < d < 4s$.
+This is smaller than the estimation rate $n^{-2s/(2s + d)}$, but larger than the minimax squared critical radius $n^{-4s/(4s + d)}$.
+As a technical matter, the problem is that when $d > 4$ there do not exist any choices of $\varepsilon$ and $K$ which satisfy both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho}, and as a result we cannot optimally balance (our upper bounds on) testing bias and variance (defined momentarily in~\eqref{eqn:testing_biasvariance}).
+Although we suspect $\varphi$ is truly suboptimal when $d > 4$, the inequality in~\eqref{eqn:testing_biasvariance} gives only an upper bound on testing bias, and thus we cannot rule out that the test $\varphi$ is optimal for all $4 < d < 4s$.
 We leave the matter to future work.
-
-That being said, it is somewhat remarkable that Laplacian eigenmaps \emph{can} take advantage of higher-order smoothness, and especially surprising that it can do so in an optimal manner.
-The sharpest known results show that the graph Laplacian eigenvectors $v_k$ converge to eigenfunctions $\psi_k$ at a rate of \textcolor{red}{(?)}.
-Naively applying these results, one can show that $\wh{f}$ to $\wt{f}$, but only at a rate far slower than the optimal rates for regression.
-Of course when the number of eigenvectors $K$ increases with $n$, as is necessary to optimally balance bias and variance of Laplacian eigenmaps, the issue only gets worse.
-Clearly, as a method for estimation and testing, the rate of convergence of Laplacian eigenmaps is much better than the rate implied by (what is currently known about) the concentration of individual eigenvectors around their continuum limits.
 
@@
 \end{equation}
-The quadratic form $\dotp{L_{n,\varepsilon}^s f_0}{f_0}_n$, eigenvalue $\lambda_{K}$, and empirical squared norm $\|f_0\|_n^2$ are each random variables that depend on the random design points $X_1,\ldots,X_n$.
-We proceed to establish suitable upper and lower bounds on these quantities.
-% (AG 8/19/21): Need to include some discussion of how the analysis contrasts with that of the classical spectral projection estimator and test.
+These design-dependent bias-variance decompositions are reminiscent of the more classical bias-variance decompositions typical in the analysis of population-level spectral projection methods, but different in certain key respects.
+Comparing~\eqref{eqn:estimation_biasvariance} and~\eqref{eqn:testing_biasvariance} to~\eqref{pf:spectral_series_estimation_3} and~\eqref{pf:spectral_series_test}, we see that two continuum objects in the latter pair of bounds, the Sobolev norm $\|f_0\|_{\mc{H}^s(\mc{X})}^2$ and the eigenvalue $\lambda_k(\Delta_P)$, have been replaced by graph-based analogues: the graph Sobolev seminorm $\dotp{L_{n,\varepsilon}^sf_0}{f_0}_n$ and the graph Laplacian eigenvalue $\lambda_k^s$.
+These latter quantities, along with the empirical squared norm $\|f_0\|_n^2$, are random variables that depend on the random design points $X_1,\ldots,X_n$.
+We proceed to establish suitable upper and lower bounds that hold in probability.
 
 \paragraph{Estimates of graph quadratic forms.}
-In Proposition~\ref{prop:graph_seminorm_fo} we restate an upper bound on the Dirichlet energy $\dotp{L_{n,\varepsilon}f}{f}_n$ from \textcolor{blue}{Green et al. 2021}. 
-\begin{proposition}[Lemma~1 of \textcolor{blue}{Green et al. 2021}]
+In Proposition~\ref{prop:graph_seminorm_fo} we restate an upper bound on the Dirichlet energy $\dotp{L_{n,\varepsilon}f}{f}_n$ from~\cite{green2021}. 
+\begin{proposition}[Lemma 1 of~\cite{green2021}]
 	\label{prop:graph_seminorm_fo}
@@
 
-In this work, we establish that an analogous bound holds for $\dotp{L_{n,\varepsilon}^sf_0}{f_0}_n$ when $s > 1$. We call this quantity the order-$s$ \emph{graph Sobolev semi-norm}.
+In this work, we establish that an analogous bound holds for $\dotp{L_{n,\varepsilon}^sf_0}{f_0}_n$ when $s > 1$. We call this quantity the \emph{graph Sobolev semi-norm}.
 \begin{proposition}
@@
 We now summarize the techniques used to prove Proposition~\ref{prop:graph_seminorm_ho}, which will help explain what role the conditions on $f_0$, $p$ and $\varepsilon$ play.
-To upper bound $\dotp{L_{n,\varepsilon}^sf}{f}_n$ in terms of $\|f\|_{H^s(\mc{X})}^2$, we introduce an intermediate quantity: the \emph{order-s non-local Sobolev seminorm} $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$.
+To upper bound $\dotp{L_{n,\varepsilon}^sf}{f}_n$ in terms of $\|f\|_{H^s(\mc{X})}^2$, we introduce an intermediate quantity: the \emph{non-local Sobolev seminorm} $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$.
 This seminorm is defined with respect to the iterated non-local Laplacian $L_{P,\varepsilon}^s = L_{P,\varepsilon} \circ \cdots \circ L_{P,\varepsilon}$, where $L_{P,\varepsilon}$ is a non-local approximation to $\Delta_P$,
@@
 In the case where $\mc{X}$ is assumed to have a boundary, the graph Laplacian $L_{n,\varepsilon}$ is a reasonable approximation of the operator $\Delta_P$ at points $x \in \mc{X}$ for which $B(x,\varepsilon) \subseteq \mc{X}$.
-In contrast, at points $x$ near the boundary of $\mc{X}$, the graph Laplacian is known to approximate a different operator altogether \citep{belkin2012}.
-This is reminiscent of the boundary effects present in the analysis of kernel smoothing.
-Thus proving convergence of $\lambda_k$ to a continuum limit becomes a substantially more challenging problem when $\mc{X}$ has a boundary.
-Rather than establishing such a result, we will instead directly use Lemma~2 of \textcolor{red}{(Green et al.
-2021)}, whose assumptions match our own, and who give a weaker bound on $\lambda_k/\lambda_k(\Delta_P)$ that will nevertheless suffice for our purposes.
-
-\begin{proposition}[Lemma~2 of \textcolor{blue}{(Green et al. 2021)}]
+In contrast, at points $x$ near the boundary of $\mc{X}$, the graph Laplacian is known to approximate a different operator altogether \citep{belkin2012}.\footnote{This is directly related to the boundary bias of kernel smoothing, since the graph Laplacian can be viewed as a kernel-based estimator of $\Delta_P$.} This renders analysis of $\lambda_k$ substantially more challenging, since its continuum limit is not $\lambda_k(\Delta_P)$.
+Rather than establishing the convergence of $\lambda_k$, we will instead use Lemma~2 of \cite{green2021}, whose assumptions match our own, and who give a weaker bound on the ratio $\lambda_k/\lambda_k(\Delta_P)$ that will nevertheless suffice for our purposes.
+
+\begin{proposition}[Lemma~2 of \cite{green2021}]
 	\label{prop:graph_eigenvalue}
@@
 \end{proposition}
-Note immediately that $\lambda_0(\Delta_P) = \lambda_0 = 0$.
-Furthermore, Weyl's Law (Lemma~\ref{lem:weyl}) tells us that under Model~\ref{def:model_flat_euclidean}, $k^{2/d} \lesssim \lambda_{k}(\Delta_P) \lesssim k^{2/d}$ for all $k \in \mathbb{N}, k > 1$.
+By our assumptions on $P$, $\lambda_0(\Delta_P) = \lambda_0 = 0$.
+Furthermore, Weyl's Law~\eqref{eqn:weyl} tells us that under Model~\ref{def:model_flat_euclidean}, $k^{2/d} \lesssim \lambda_{k}(\Delta_P) \lesssim k^{2/d}$ for all $k \in \mathbb{N}, k > 1$.
 Combining these statements with~\eqref{eqn:graph_eigenvalue}, we conclude that $\lambda_{K} = \Omega_P(K^{2/d})$ so long as $K \lesssim \varepsilon^{-d}$.
@@
 Finally, in order to show that $\varphi$ has small Type II error whenever~$\|f_0\|_P$ is greater than the critical radius given by~\eqref{eqn:sobolev_space_testing_critical_radius}, we require a lower bound on $\|f_0\|_n^2$ in terms of $\|f_0\|_P^2$.
-In Proposition~\ref{prop:empirical_norm_sobolev} we establish that such a one-sided bound holds, whenever $\|f_0\|_P$ is sufficiently large.
+In Proposition~\ref{prop:empirical_norm_sobolev} we establish that such a one-sided bound holds whenever $\|f_0\|_P$ is sufficiently large.
 \begin{proposition}
@@
 Finally, invoking the bounds of Propositions~\ref{prop:graph_seminorm_fo}-\ref{prop:empirical_norm_sobolev} inside the bias-variance tradeoffs~\eqref{eqn:estimation_biasvariance} and~\eqref{eqn:testing_biasvariance} and then choosing $K$ to balance bias and variance (when possible), leads to the conclusions of Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}.
+
+% (AG 8/27/21): Following chunk pasted here temporarily.
+On the other hand, it is somewhat remarkable that Laplacian eigenmaps can \emph{ever} take advantage of higher-order smoothness in $f_0$.
+The sharpest known results show that the graph Laplacian eigenvectors $v_k$ converge to eigenfunctions $\psi_k$ at a rate of \textcolor{red}{(?)}.
+Naively applying these results, one can show that $\wh{f}$ to $\wt{f}$, but only at a rate far slower than the optimal rates for regression.
+Of course when the number of eigenvectors $K$ increases with $n$, as is necessary to optimally balance bias and variance of Laplacian eigenmaps, the issue only gets worse.
+Clearly, as a method for estimation and testing, the rate of convergence of Laplacian eigenmaps is much better than the rate implied by (what is currently known about) the concentration of individual eigenvectors around their continuum limits.
 
@@
 As discussed previously, when $\mc{X}$ is a domain without boundary and $\Delta_P$ is the manifold weighted Laplace-Beltrami operator, appropriate bounds on the graph eigenvalues $\lambda_k$ have already been derived in \citep{burago2014,trillos2019,garciatrillos19}.
-The precise result we need is a simple consequence of Theorem 2.4 of~\citep{calder2019}.
+The precise result we need is a direct consequence of Theorem 2.4 of~\citep{calder2019}.
 \begin{proposition}[\textbf{c.f Theorem 2.4 of~\citep{calder2019}}]
@@
 \end{proposition}
-(For the specific computation used to deduce Proposition~\ref{prop:graph_eigenvalue_manifold} from Theorem 2.4 of~\citep{calder2019}, see~\textcolor{blue}{(Green 2021)}.)
+(For the specific computation used to deduce Proposition~\ref{prop:graph_eigenvalue_manifold} from Theorem 2.4 of~\citep{calder2019}, see~\cite{green2021}.)
 
@@
 \end{proposition}
-We prove Proposition~\ref{prop:empirical_norm_sobolev_manifold} in a parallel manner to its flat Euclidean counterpart (Proposition~\ref{prop:empirical_norm_sobolev}), by using a Gagliardo-Nirenberg inequality for functions on compact Riemannian manifolds, and then applying a one-sided Bernstein's inequality.
+We prove Proposition~\ref{prop:empirical_norm_sobolev_manifold} in a parallel manner to its flat Euclidean counterpart (Proposition~\ref{prop:empirical_norm_sobolev}), by first using a Gagliardo-Nirenberg inequality to upper bound the $L^4(\mc{X})$ norm of a Sobolev function defined on a compact Riemannian manifold, and then applying a one-sided Bernstein's inequality.
 Finally, combining Propositions~\ref{prop:graph_seminorm_manifold}-\ref{prop:empirical_norm_sobolev_manifold} with the conditional-on-design bias-variance decompositions~\eqref{eqn:estimation_biasvariance} and \eqref{eqn:testing_biasvariance} leads to the conclusions of Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold}.

### 2021-08-28 manuscripts/laplacian_eigenmaps/sections/introduction.tex
Through upper bounds on Laplacian eigenmaps in flat Euclidean case.

@@
 Together, these insights give a reasonably complete explanation of how graph spectra \emph{learn} the structure of an underlying distribution.
-In contrast, in this paper we develop theory showing that graph spectral methods can effectively \emph{leverage} structure to perform downstream (supervised) learning tasks.
+In contrast, in this paper we develop theory showing that graph spectral methods can effectively \emph{leverage} learned structure to perform downstream (supervised) learning tasks.
 Our main goal is to show that spectral projection procedures--which use eigenvectors of a neighborhood graph Laplacian as features in a linear regression problem---are statistically optimal methods for two problems in nonparametric regression: estimation and goodness-of-fit testing.
@@
 Given samples $(X_1,Y_1),\ldots,(X_n,Y_n)$, Laplacian eigenmaps first forms a neighborhood graph, with vertices at the design points $X_1,\ldots,X_n$ and weighted edges between pairs of sufficiently close points $X_i$ and $X_j$.
-The Laplacian eigenmaps estimator $\wh{f}$ is then computed using the spectral decomposition of an (unweighted) Laplacian defined over this neighborhood graph, by projecting the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto a subspace spanned by a certain number of ``low-frequency'' graph Laplacian eigenvectors; more precisely, the eigenvectors corresponding to the smallest eigenvalues of the Laplacian.
+The Laplacian eigenmaps estimator $\wh{f}$ is then computed using the spectral decomposition of an (unweighted) Laplacian defined over this neighborhood graph, by projecting the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto a subspace spanned by a certain number of ``low-frequency'' graph Laplacian eigenvectors; more precisely, these are the eigenvectors corresponding to the smallest eigenvalues of the graph Laplacian.
 The Laplacian eigenmaps test statistic $\wh{T} = \|\wh{f}\|_n^2$ can in turn be used as a statistic to detect whether any signal is present, that is, whether the regression function $f_0(x) = \mathbb{E}[Y|X = x]$ is anywhere non-zero.
@@
 Of course, this comes with a tradeoff.
-Intuitively, using an empirical approximation to the ``right'' population-level basis will incur some additional error; the question is whether, notwithstanding this additional error, Laplacian eigenmaps shares the strong statistical properties of classical spectral projection.
+Intuitively, using an empirical approximation to the ``right'' population-level basis will incur some additional error.
+The question is whether, notwithstanding this additional error, Laplacian eigenmaps shares the strong statistical properties of classical spectral projection.
 
@@
 	\item Over the Sobolev space $H^{s}(\mc{X})$, we establish the Laplacian eigenmaps estimator $\wh{f}$ has in-sample mean-square error on the order of $n^{-2s/(2s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d$.
-A test based on the statistic $\|\wh{f}\|_n^2$ has a critical radius on the order of $n^{-4s/(4s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d \in \{1,2,3,4\}$.
+A test based on the statistic $\|\wh{f}\|_n^2$ has a squared critical radius on the order of $n^{-4s/(4s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d \in \{1,2,3,4\}$.
 	\item The estimator $\wh{f}$ is defined only in-sample.
@@
 \end{itemize}
-In all cases, our bounds also depend optimally on the radius $M$ of the Sobolev ball under consideration.
+% (AG 8/27/21): Removed the following sentence, because I no longer know if it is true.
+% In all cases, our bounds also depend optimally on the radius $M$ of the Sobolev ball under consideration.
 
-To derive these results, we develop an analysis framework reminiscent of that used in the analysis of classical spectral projection estimators, but ultimately quite different.
-The chief difference is in the approximation error (bias): for our purposes we must establish that with high probability, the eigenvectors of graph Laplacians (which are data-dependent and hence random objects) effectively approximate continuum Sobolev functions, whereas in the classical case an upper bound on the approximation error follows immediately from the aforementioned spectral characterization of Sobolev spaces.
+To derive these results, we develop an framework for analysis reminiscent of that  classically used to analyze population-level spectral projection estimators, but ultimately quite different.
+The chief difference is in the approximation error (bias).
+For our purposes we must establish that with high probability, the eigenvectors of graph Laplacians (which are data-dependent and hence random objects) effectively approximate continuum Sobolev functions, whereas in the classical analysis an upper bound on the approximation error follows immediately from the aforementioned spectral characterization of Sobolev spaces.
 To tightly upper bound the approximation error of Laplacian eigenmaps, we rely on several recent results regarding concentration of the spectra of graph Laplacians around their population limits, and prove some additional results which may be of independent interest.
@@
 
-In the random design setting the analysis is more challenging, and only a few works have studied the matter.
+In the random design setting the analysis of graph-based approaches to supervised learning is more challenging, and only a few works have studied the matter.
 \cite{zhou2011} provide an asymptotic analysis of the Laplacian eigenmaps estimator in the semi-supervised setting, as the number of unlabeled samples tends to infinity; in this case the estimator reduces to the classical spectral projection estimator, and there is no extra error incurred by using data-dependent eigenvectors.

### 2021-08-28 manuscripts/laplacian_eigenmaps/sections/preliminaries.tex
Through upper bounds on Laplacian eigenmaps in flat Euclidean case.

@@
 \end{definition}
-Boundary conditions plays an important role in the analysis of spectral methods, as we explain further in Section~\ref{subsec:spectral_projection}.
+Boundary conditions play an important role in the analysis of spectral methods, as we explain further in Section~\ref{subsec:spectral_projection}.
 For now, we limit ourselves to pointing out that for functions $f \in C^\infty(\mc{X})$, the zero-trace condition can be stated more concretely, as implying that $\partial^{k}f/\partial{\bf n}^k(x) = 0$ for each $k = 0,\ldots,s - 1$, and for all $x \in \partial\mc{X}$.
@@
 
-\subsection{Classical Spectral Projection}
+\subsection{Spectral Projection using Continuum Eigenfunctions}
 \label{subsec:spectral_projection}
 Laplacian Eigenmaps serves as a data-dependent alternative to more classical spectral projection methods.
-Classically, spectral projection methods operate by projecting the response vector ${\bf Y}$ onto a subspace spanned by eigenfunctions of some differential operator; for instance, the Laplacian operator $\Delta = \sum_{i = 1}^{d} \partial^2f/\partial x_i^2$, which corresponds to projection onto a Fourier basis.
-We will focus on a density-dependent analog to the Laplacian,  $\Delta_P:C^2(\mc{X}) \to L^2(\mc{X})$,
+Classically, spectral projection methods operate by projecting the response vector ${\bf Y}$ onto a subspace spanned by eigenfunctions of some differential operator.
+For instance, projecting ${\bf Y}$ onto the eigenfunctions of the Laplacian operator $\Delta = \sum_{i = 1}^{d} \partial^2f/\partial x_i^2$ corresponds to estimating $f_0$ using noisy empirical Fourier coefficients.
+
+We will focus on a density-dependent altnerative to the Laplacian,  $\Delta_P:C^2(\mc{X}) \to L^2(\mc{X})$,
 \begin{equation}
@@
 \end{equation}
-and sorted as usual in ascending order of eigenvalue.\footnote{For formal justification of why~\eqref{eqn:laplace_beltrami_eigenproblem} has discrete spectrum under either Model~\ref{def:model_flat_euclidean} or~\ref{def:model_manifold}, see~\cite{garciatrillos18,trillos2019}.} The classical spectral projection estimator and test statistic are then simply
+and sorted as usual in ascending order of eigenvalue.\footnote{For formal justification of why~\eqref{eqn:laplace_beltrami_eigenproblem} has discrete spectrum under either Model~\ref{def:model_flat_euclidean} or~\ref{def:model_manifold}, see~\cite{garciatrillos18,trillos2019}.} Formally, the classical spectral series estimator and test statistic are\footnote{Technically speaking, $\wt{f}$ is not the projection of ${\bf Y}$ onto the subspace of $L^2(P_n)$ spanned by $\psi_1,\ldots,\psi_K$, since $\dotp{\psi_k}{\psi_{\ell}}_n \neq 0$.
+However, these eigenfunctions are very close to orthogonal in $L^2(P_n)$, orthogonalization will have a negligible impact on the overall statistical properties, and so for simplicity we consider the methods as defined in~\eqref{eqn:classical_spectral_projection}.}
 \begin{equation}
@@
 \end{equation}
+
 We focus on the particular differential operator $\Delta_P$ because the spectrum of $\Delta_P$ is the population-level limit of graph Laplacian spectra: for any fixed $k \in \mathbb{N}$, as the number of samples $n \to \infty$ and the graph radius $\varepsilon \to 0$,
@@
 
-On the other hand, the eigenvectors and eigenfunctions of $\Delta_P$ can also be used to define Sobolev spaces in an alternative manner. Consider the ellipsoid
+On the other hand, the eigenvectors and eigenfunctions of $\Delta_P$ can also be used to give a spectral definition of Sobolev spaces. Consider the ellipsoid
 \begin{equation}
@@
 Suppose additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$, $f_0 \in \mc{H}^{s}(\mc{X};M)$ and $\|f_0\|_P^2 \leq 1$.
-Then there exists a constant $C$ which does not depend on $f_0,M$ or $n$ such that the following statement holds: if the spectral projection estimator $\wt{f}$ is computed with parameter $K = \floor{M^2n}^{d/(2s + d)} \vee 1$, then
+Then there exists a constant $C$ which does not depend on $f_0,M$ or $n$ such that the following statement holds: if the spectral projection estimator $\wt{f}$ is computed with parameter $K = \floor{M^2n}^{d/(2s + d)}$, then
 	\begin{equation}
 	\label{eqn:spectral_series_estimation}
-	\Ebb\bigl[\|\wt{f} - f_0\|_P^2\bigr] \leq C \min\bigl\{M^2(M^2n)^{-2s/(2s + d)}, M^2\bigr\}.
+	\Ebb\bigl[\|\wt{f} - f_0\|_P^2\bigr] \leq C \min\Bigl\{M^2\bigl(M^2n\bigr)^{-2s/(2s + d)}, M^2\Bigr\}.
 	\end{equation}
 \end{proposition}
-When the Sobolev ball radius $M \asymp 1$, the upper bound in~\eqref{eqn:spectral_series_estimation} is on the order of $n^{-2s/(2s + d)}$, which matches the standard minimax estimation rate for Sobolev classes (see~\cite{wasserman2006,tsybakov08} and references therein).
+When the Sobolev ball radius $M \asymp 1$---which automatically implies $\|f_0\|_{P}^2 \lesssim 1$ as well---the upper bound in~\eqref{eqn:spectral_series_estimation} is on the order of $n^{-2s/(2s + d)}$, which matches the well-known minimax rate for estimation over Sobolev classes (see e.g.~\cite{wasserman2006,tsybakov08} and references therein).
 
@@
 We now give the proof of Proposition~\ref{prop:spectral_series_estimation}.
-The structure of the analysis, which is fairly classical and straightforward, can be usefully compared and contrasted with our analysis of Laplacian Eigenmaps (see Section~\ref{subsec:analysis}).
+The structure of the analysis, which is fairly classical and straightforward, can be usefully compared to our analysis of Laplacian Eigenmaps (see Section~\ref{subsec:analysis}).
 
@@
 \end{equation}
-The claim of the proposition then follows from two (variants of) classical results in spectral geometry.
+The claim of the proposition then follows from variants of two classical results in spectral geometry.
 The first is a Weyl's Law asymptotic scaling of the eigenvalues of $\Delta_P$ due to~\cite{dunlop2020}; formally, there exist constants $c$ and $C$ (which will depend on $P$ and $d$) such that
@@
 \end{equation}
-The second is a local analog to Weyl's Law,
+The second is a local analog to Weyl's Law, which says that there exists a constant $C$ (again depending on $P$ and $d$) such that
 \begin{equation}
 \label{eqn:local_weyl}
-\sup_{x \in \mc{X}}\biggl\{\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2\biggr\} \leq CK \quad\textrm{for all $K \in \mathbb{N}$},
+\sup_{x \in \mc{X}}\biggl\{\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2\biggr\} \leq CK \quad\textrm{for all $K \in \mathbb{N}$}.
 \end{equation}
-which follows straightforwardly from Theorem 17.5.3 of~\cite{hormander1973}.
-Plugging the upper bounds~\eqref{eqn:weyl} and~\eqref{eqn:local_weyl} back into~\eqref{pf:spectral_series_estimation_1} and applying the assumed upper bound $\Ebb[(f_0(X))^2] \leq 1$ shows that
-\begin{equation*}
-\Ebb \|\wt{f} - f_0\|_P^2 \leq C\biggl(\frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{(k + 1)^s} + \frac{K}{n}\biggr),
-\end{equation*}
-and setting $K = \floor{M^2n}^{d/(2s + d)} \vee 1$, which balances squared bias and variance, then yields the claim.
+Equation~\eqref{eqn:local_weyl} follows straightforwardly from Theorem 17.5.3 of~\cite{hormander1973}.
+Plugging the upper bounds~\eqref{eqn:weyl} and~\eqref{eqn:local_weyl} back into~\eqref{pf:spectral_series_estimation_1} shows that
+\begin{equation}
+\label{pf:spectral_series_estimation_3}
+\Ebb \|\wt{f} - f_0\|_P^2 \leq C\biggl(\frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{(K + 1)^{2s/d}} + \frac{K}{n}\biggr),
+\end{equation}
+and setting $K = \floor{(1 + \|f_0\|_{P}^2)M^2n}^{d/(2s + d)}$, which balances squared bias and variance, then yields the claim.
 \qed.
@@
 \end{equation} 
-To fix ideas, here and throughout we focus on the signal detection problem, which is the special case of $f_0^{\star} = 0$.\footnote{This is without loss of generality since all the test statistics we consider are easily modified to handle the case when $f_0^{\ast}$ is not $0$, by simply subtracting $f_0^{\ast}(X_i)$ from each observation $Y_i$, with no change in the analysis.}
+To fix ideas, here and throughout we focus on the signal detection problem, meaning the special case where $f_0^{\star} = 0$.\footnote{This is without loss of generality since all the test statistics we consider are easily modified to handle the case when $f_0^{\ast}$ is not $0$, by simply subtracting $f_0^{\ast}(X_i)$ from each observation $Y_i$, with no change in the analysis.}
 
-In this case, the classical spectral projection test $\wt{\varphi} = \1\{\wt{T} \geq K/N + \sqrt{K/a}/n\}$ has bounded Type I error, $\Ebb_{0}[\wt{\varphi}] \leq a$.
-Proposition~\ref{prop:spectral_series_testing} gives an upper bound on the Type II error that holds uniformly over all $f_0 \in \mc{H}^s(\mc{X};M)$ which have a sufficiently large $L^2(P)$ norm.
+In this case, the classical spectral projection test $\wt{\varphi} = \1\{\wt{T} \geq K/N + \sqrt{2K/an}\}$ has bounded Type I error, $\Ebb_{0}[\wt{\varphi}] \leq a (1 + o(1))$.
+Proposition~\ref{prop:spectral_series_testing} gives an upper bound on the Type II error that holds uniformly over all $f_0 \in \mc{H}^s(\mc{X};M)$ for which $\|f_0\|_P^2$ is sufficiently large.
 \begin{proposition}
@@
 	Suppose data is observed according to Model~\ref{def:model_flat_euclidean}, and that the density $p$ is known.
-Suppose additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$ and $f_0 \in \mc{H}^{s}(\mc{X};M)$.
+Suppose additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$, $f_0 \in \mc{H}^{s}(\mc{X};M)$ and $\|f_0\|_{\Leb^4(\mc{X})}^4 \leq 1$.
 Then there exists a constant $C$ which does not depend on $f_0,M$ or $n$ such that the following statement holds: if the spectral projection test $\wt{\varphi}$ is computed with parameter $K = \floor{M^2n}^{2d/(4s + d)}$, and if
@@
 	\label{eqn:spectral_series_testing}
-	\|f_0\|_P^2 \geq \textcolor{red}{(?)}
+	\|f_0\|_P^2 \geq C\min\{M^2(M^2n)^{-4s/(4s + d)}, M^2\}
 	\end{equation}
@@
 \end{proposition}
-Again, when the Sobolev ball radius $M \asymp 1$, the upper bound in~\eqref{eqn:spectral_series_testing} becomes $n^{-4s/(4s + d)}$, matching the usual minimax critical radius over Sobolev spaces (for more details on goodness-of-fit testing see~\citep{ingster2009,ingster2012}).
-% (AG 8/19/21): Same comment as above.
+Focusing again on the special case where the Sobolev ball radius $M \asymp 1$---which automatically implies $\|f_0\|_{\Leb^4(\mc{X})}^4 \lesssim 1$ when $s > d/4$, due to a Sobolev embedding theorem---the upper bound in~\eqref{eqn:spectral_series_testing} is $n^{-4s/(4s + d)}$, matching the usual minimax critical radius over Sobolev spaces (for more details on goodness-of-fit testing see~\cite{ingster2009,ingster2012}).
+On the other hand, when $s \leq d/4$ the problem fundamentally changes in two ways.
+First, Sobolev space $H^s(\mc{X})$ no longer continuously embeds into $L^4(\mc{X})$, and indeed the minimax testing rates in this case are unknown.
+Second, if one explicitly assumes $f_0 \in L^4(\mc{X};1)$ then the minimax critical radius for regression testing is on the order of $n^{-1/2}$ \citep{guerre02}, and is achieved by a test using the naive statistic $\|{\bf Y}\|_n^2$.
+In other words, the regression testing problem over Sobolev spaces fundamentally changes when $s \leq d/4$, and hereafter when we discuss testing we will limit our consideration to $s > d/4$.
+\paragraph{Proof of Proposition~\ref{prop:spectral_series_testing}.}
+We briefly lay out the main ideas needed to prove Proposition~\ref{prop:spectral_series_testing}, following the lead of~\cite{ingster2009} who prove a similar result in the special case where $P$ is the uniform distribution over $\mc{X} = [0,1]^d$, and referring to that work for more details.
 
-The main takeaway from Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} is that spectral projection methods achieve optimal rates over Sobolev classes, when the design density $p$ satisfies only a nonparametric notion of smoothness\footnote{The assumption $p \in C^{\infty}(\mc{X})$ could likely be weakened.
-Alternatively, one could likely derive upper bounds assuming Model~\ref{def:model_manifold} which depend only on the intrinsic dimension $m$.
-Since these would not substantially add to the main points of Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, we do not pursue the details further.} but, crucially, is known a priori.
-As we will see, Laplacian eigenmaps achieves similar rates of convergence when $p$ is unknown, and in this sense both learns and leverages the design distribution in a way that more classical spectral projection methods cannot.
-It is worth pointing out that other methods besides Laplacian eigenmaps are statistically optimal for nonparametric regression even when $p$ is unknown.
+We begin by computing the first two moments of the test statistic $\wt{T}$. The expectation is
+\begin{align*}
+\Ebb[\wt{T}] = \frac{(n - 1)}{n} \sum_{k = 1}^{K} \dotp{f_0}{\psi_K}_P^2 + \frac{K}{n} + \Ebb[(f_0(X))^2 \dot \sum_{k = 1}^{K} (\psi_k(X))^2],
+\end{align*}
+from which we clearly see that under the alternative,
+\begin{equation*}
+\Ebb_{f_0}[\wt{T}] \geq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\lambda_{K + 1}(\Delta_{P})^s} + K/n.
+\end{equation*}
+To compute the variance, we decompose $\wt{T} = \wt{T}_{1,1} + \wt{T}_{1,2} + \wt{T}_{1,3} + \wt{T}_2$ into the sum of 3 U-statistics and the remaining diagonal terms; letting $\kappa(x,x') = \sum_{k = 1}^{K} \psi_k(x) \psi_k(x')$ be the equivalent kernel, we have that
+\begin{align*}
+T_{1,1} & = \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} w_i w_j \kappa(X_i,X_j),\quad && T_{1,2} = \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} \bigl(w_i f_0(X_j) + w_jf_0(X_i)\bigr)\kappa(X_i,X_j) \\
+T_{1,3} & = \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} f_0(X_i) f_0(X_j) \kappa(X_i,X_j), \quad && T_2 = \frac{1}{n^2} \sum_{i = 1}^{n} Y_i^2 \kappa(X_i,X_i).
+\end{align*}
+The variances of each statistic can be found by routine computation (see~\cite{ingster2009}), and in particular satisfy the upper bounds
+\begin{align*}
+\Var(T_{1,1}) & \leq \frac{2K}{n^2}, \quad && \Var(T_{1,2}) \overset{\mathrm{(i)}}{\leq} \frac{C}{n}\|f_0\|_P^2\\
+\Var(T_{1,3}) & \overset{\mathrm{(ii)}}{\leq} C\biggl(\frac{K}{n}\|f_0\|_P^4 + \frac{K}{n^2}\|f_0\|_{L^4(\mc{X})}^4\biggr), \quad && \Var(T_{2}) \overset{\mathrm{(iii)}}{\leq} \frac{CK^2}{n^3}\biggl(1 + \|f_0\|_{L^4(\mc{X})}^4\biggr)
+\end{align*}
+where $\mathrm{(i)}-\mathrm{(iii)}$ hold due to local Weyl's law, meaning the upper bound on $\sum_{k = 1}^{K} (\psi_k(x))^2$ given in~\eqref{eqn:local_weyl}. Upper bounds on Type I and Type II error,
+\begin{equation*}
+\Ebb_0[\wt{\varphi}] \leq \biggl(1 + CK/n^2\biggr)a, \quad \Ebb_{f_0}[1 - \wt{\varphi}] \leq \frac{C(K/n^2 + \|f_0\|_P^2 + K/n \|f_0\|_P^4 + K/n^2 \|f_0\|_{L^4(\mc{X})}^4)}{(\sum_{k = 1}^{K} \dotp{f_0}{\psi_k}_P^2 - \sqrt{2K/an})^2},
+\end{equation*}
+follow from Chebyshev's inequality. It can be verified that so long as
+\begin{equation}
+\label{pf:spectral_series_test}
+\|f_0\|_P^2 \geq C\Biggl(\frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\lambda_{K + 1}(\Delta_{P})^s} + \sqrt{\frac{2K}{n}}\biggl(\sqrt{\frac{1}{a}} + \sqrt{\frac{1}{b}}\biggr)\Biggr)
+\end{equation}
+for a sufficiently large constant $C$, then $\Ebb_{f_0}[1 - \wt{\varphi}] \leq b$.
+The two summands in~\eqref{pf:spectral_series_test} are bias and standard deviation terms, respectively; choosing $K = \floor{M^2n}^{2d/(4s + d)}$ balances these two terms, and leads to~\eqref{eqn:spectral_series_testing}.
+\qed
+% (AG 8/19/21): Same comment as above, regarding the role of M in random design regression testing.
+
+The main takeaway from Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} is that spectral projection methods for nonparametric regression achieve optimal rates of convergence, when the regression function $f_0$ is Sobolev smooth and the design density $p$ is known a priori and satisfies an appropriate notion of smoothness\footnote{The assumption $p \in C^{\infty}(\mc{X})$ could likely be weakened, but since this would not substantially add to the main points of Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, we do not pursue the details further.}.
+As we will see, Laplacian eigenmaps achieves comparable rates of convergence when $p$ is sufficiently smooth but potentially unknown, and in this sense both learns and leverages the design distribution in a way that more classical spectral projection methods do not.
+
+Of course, it is worth pointing out that other methods besides Laplacian eigenmaps are statistically optimal for nonparametric regression even when $p$ is unknown.
 We comment more on some of these in Section~\ref{sec:discussion}, after we have derived our major results regarding Laplacian Eigenmaps.
-
-\paragraph{In-sample mean squared error.}
-As mentioned in our introduction, roughly speaking one of our main conclusions is that the Laplacian eigenmaps estimator $\wh{f}$ is minimax rate-optimal.
-It is worth being clear about what we do and do not mean by this statement.
-We do not mean that the estimator $\wh{f}$ will match the upper bound given in~\eqref{eqn:sobolev_space_minimax_estimation_rate}, since such a statement does not make sense when the estimator is defined only at the random design points $X_1,\ldots,X_n$.
-Instead we will measure loss using the squared $L^2(P_n)$ error.
-In Section~\ref{sec:out_of_sample} we show that an extension of $\wh{f}$ defined over all $\mc{X}$ has $L^2(P)$ error comparable to the $L^2(P_n)$ error of $\wh{f}$.
-We also believe that in the random design setting we work in, simple arguments will imply that $L^2(P_n)$ risk has the same minimax rate of convergence as $L^2(P)$ risk.
-We sketch such an argument in Section~\ref{sec:out_of_sample}, but do not further pursue the details.
-
-% (SB via AG): Move this somewhere later.
-Additionally, we will not actually measure accuracy using the expectation of the loss.
-Rather, we will give a constant probability bound on $\|\cdot\|_n^2$.
-For instance, when $f_0 \in H^1(\mc{X};1)$, we will show that with probability $1 - \delta$ the loss $\|\wh{f} - f_0\|_n^2 \leq C_{\delta} n^{-2/(2 + d)}$, for a constant $C_{\delta}$ that depends on $\delta$ but not on $f_0$ or $n$.
-Thus we give an upper bound on the $(1 - \delta)$th quantile of $\|\wh{f} - f_0\|_n^2$, rather than an upper bound on its expectation.
-We explain the reason for this in Section~\ref{sec:minimax_optimal_laplacian_eigenmaps}.
-We also show that if $f_0$ is bounded in a larger norm---for instance, if it is H\"{o}lder rather than Sobolev smooth---then we can obtain bounds on the expected $L^2(P_n)$ loss.
-
-There is one other subtlety introduced by the use of in-sample mean squared error.
-Technically speaking, elements $f \in H^s(\mc{X})$ are equivalence classes, defined only up to a set of measure zero.
-Thus one cannot speak of the pointwise evaluation $f_0(X_i)$, as we do by defining our target of estimation to be $f_0(X_i)$, $i=1,\ldots,n$, until one selects \emph{representatives}.
-When $s > d/2$, every element $f$ of $H^s(\mc{X})$ admits a continuous version $f^{\ast}$, and as is standard we set this to be our favored representative.
-When $s \leq d/2$, some elements in $H^s(\mc{X})$ do not have any continuous version; however they admit a \emph{quasi-continuous} version \citep{evans15} known as the \emph{precise representative}, and we use this representative.
-To be clear, however, it does not really matter which representative we choose.
-Since all versions agree except on a set of measure zero, and since $P$ is absolutely continuous with respect to Lebesgue measure (in Model~\ref{def:model_flat_euclidean}) or the volume form $d\mu$ (in Model~\ref{def:model_manifold}), with probability $1$ any two versions $g_0, h_0 \in f_0$ will satisfy $g_0(X_i) = h_0(X_i)$ for all $i = 1,\ldots,n$.
-The bottom line is that we can use the notation $f_0(X_i)$ without fear of ambiguity or confusion.
-
-Finally, we note that for testing none of these comments are relevant.
-We will show that our test has small worst-case risk whenever $\epsilon \gtrsim \epsilon_n(H_0^s(\mc{X};M))$, thus establishing that it is a minimax optimal test in the usual sense.

### 2021-08-29 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Partitioned off discussion section.

@@
 \paragraph{Empirical norm.}
-Finally, in order to show that $\varphi$ has small Type II error whenever~$\|f_0\|_P$ is greater than the critical radius given by~\eqref{eqn:sobolev_space_testing_critical_radius}, we require a lower bound on $\|f_0\|_n^2$ in terms of $\|f_0\|_P^2$.
-In Proposition~\ref{prop:empirical_norm_sobolev} we establish that such a one-sided bound holds whenever $\|f_0\|_P$ is sufficiently large.
+Finally, in Proposition~\ref{prop:empirical_norm_sobolev} we establish that a one-sided bound of the form $\|f_0\|_n^2 \gtrsim \|f_0\|_P^2$ whenever $\|f_0\|_P^2$ is itself sufficiently large.
 \begin{proposition}
@@
 \end{proposition}
-To prove Proposition~\ref{prop:empirical_norm_sobolev}, we use the Gagliardo-Nirenberg interpolation inequality (see e.g.
-Theorem~(?) of~\citep{evans10}) to control the $4$th moment of $f$ in terms of $\|f\|_P$ and $|f|_{H^s(\mc{X})}$, then invoke a one-sided Bernstein's inequality as in \cite[Section 14.2]{wainwright2019}.
-Note carefully that the statement~\eqref{eqn:empirical_norm_sobolev} is \emph{not} a uniform guarantee over all $f \in H^s(\mc{X};M)$, as such a statement cannot hold in the sub-critical regime ($2s \leq d$).
-Fortunately, a pointwise bound---meaning a bound that holds with high probability for a single $f \in H^s(\mc{X})$---is sufficient for our purposes.
-% (AG) Ask Ryan for reference to Evans.
-% (SB via AG): More discussion about the uniform guarantee point.
+To prove Proposition~\ref{prop:empirical_norm_sobolev}, we use a Gagliardo-Nirenberg interpolation inequality (see e.g.
+Theorem~12.83 of~\citep{leoni2017}) to control the $4$th moment of $f \in H^s(\mc{X})$ in terms of $\|f\|_P$ and $|f|_{H^s(\mc{X})}$, then invoke a one-sided Bernstein's inequality as in \cite[Section 14.2]{wainwright2019}.
+Note carefully that the statement~\eqref{eqn:empirical_norm_sobolev} is \emph{not} a uniform guarantee over all $f \in H^s(\mc{X};M)$.
+Indeed, such a statement cannot hold in the sub-critical regime ($2s \leq d$).\footnote{This is because in the sub-critical regime, for any set of points $\{x_1,\ldots,x_n\}$ there exist a sequence of functions $(f_k)_{k \in \mathbb{N}} \subset H^s(\mc{X};1)$ satisfying $f_k(x_i) = 1$ for each $i = 1,\ldots,n$, but for which $\|f_k\|_P^2 \to 0$ as $k \to \infty$.} Fortunately, a pointwise bound---meaning a bound that holds with high probability for a single $f \in H^s(\mc{X})$---is sufficient for our purposes.
 
@@
 % (AG 8/27/21): Following chunk pasted here temporarily.
-On the other hand, it is somewhat remarkable that Laplacian eigenmaps can \emph{ever} take advantage of higher-order smoothness in $f_0$.
-The sharpest known results show that the graph Laplacian eigenvectors $v_k$ converge to eigenfunctions $\psi_k$ at a rate of \textcolor{red}{(?)}.
+We conclude this section by pointing out that it is somewhat remarkable how Laplacian eigenmaps can take advantage of higher-order smoothness in $f_0$.
+The sharpest known results \citep{calder2019,cheng2021} show that the squared error $\|v_k - \psi_k\|_n^2$ converges to $0$ at a rate of $n^{-2/(4 + d)}$).
 Naively applying these results, one can show that $\wh{f}$ to $\wt{f}$, but only at a rate far slower than the optimal rates for regression.
@@
 
-	We now explain this discrepancy, between the full-dimensional and manifold settings.
+	We now explain this discrepancy between our rates in the flat Euclidean and manifold settings.
 At a high level, thinking of the graph $G$ as an estimate of the manifold $\mc{X}$, we incur some error in this estimate by using Euclidean distance rather than geodesic distance to form the edges of $G$.
@@
 
-\section{Discussion}
-\label{sec:discussion}
-
-\subsection{Comparison with other estimators}
-In this paper, we have motivated Laplacian eigenmaps by viewing it as a noisy approximation of a classical spectral projection method, which is its most obvious counterpart.
-We have shown that Laplacian eigenmaps inherits the optimality properties of its more classical counterpart.
-We now discuss the relationship between Laplacian eigenmaps and three other approaches to nonparametric regression.
-The first two---nonparametric least squares and kernel smoothing---are classical, whereas the third---Laplacian smoothing---makes use of the graph Laplacian in a different way than Laplacian eigenmaps.
-
-% (SB via AG): Change the name of this to ``spectral projection'' or something like that.
-\paragraph{Nonparametric least squares.}
-The standard recommended alternative to spectral projection methods, when the distribution $P$ is considered non-uniform or unknown, is to do least-squares.
-For example, suppose instead of knowing $\psi_1,\psi_2,\ldots$, we had access only to eigenfunctions $\phi_1,\phi_2,\ldots$ of an unweighted Laplace-Beltrami operator $\Delta$.
-Then letting $\Phi_K = \mathrm{span}\{\phi_1,\ldots,\phi_K\}$, the least-squares estimator and test statistic
-\begin{equation*}
-\wt{f}_{\mathrm{LS}} := \argmin_{f \in \Phi_K} \|Y - f\|_n^2,\textrm{and}~~\wt{T}_{\mathrm{LS}} = \|\wt{f}\|_{\nu}^2
-\end{equation*}
-are still rate-optimal over $H_0^s(\mc{X})$.
-This holds true for both Model~\ref{def:model_flat_euclidean} and~\ref{def:model_manifold}, and is a consequence of our assumption that $p$ is bounded away from $0$.
-
-However, this is not a totally satisfactory fix.
-For one thing, the least squares approach just outlined requires that we know the domain $\mc{X}$, in the strong sense that we know the Laplace-Beltrami operator $\Delta$ defined on $\mc{X}$.
-Domain knowledge is generally precious information, and such strong knowledge of $\mc{X}$ seems particularly unrealistic in the case where $\mc{X}$ is a manifold, and $\Delta$ is the manifold Laplace-Beltrami operator.
-Additionally, even if we know $\mc{X}$, diagonalizing the Laplace-Beltrami operator $\Delta$ is quite difficult for all but a few special domains, such as the unit cube $\mc{X} = [0,1]^d$ or torus $\mc{X} = \mathbb{T}^d$.
-
-% (SB via AG): Maybe also worth discussing the ill-posed counterpart?
-
-\paragraph{Kernel smoothing.}
-It is also natural to ask whether the two stage estimator $T_{n,h}\wh{f}$ defined in Section~\ref{sec:out_of_sample} has any advantage over the simpler approach of directly kernel smoothing the responses, i.e.
-using the estimator $T_{n,h}Y$ (possibly for a different choice of $h$).
-In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{n,h}\wh{f}\|_P^2$ is of a strictly lower order than $\inf_{h'} \Ebb\|f_0 - T_{h',n}Y\|_P^2$.
-This is possible because Laplacian eigenmaps induces a completely different bias than kernel smoothing.
-For example, when $f_0$ and $p$ satisfy the so-called \emph{cluster} assumption--- e.g.
-$f_0$ is piecewise constant in high-density regions (clusters) of $p$--- then the bias of Laplacian eigenmaps can much smaller than that of kernel smoothing (for equivalent levels of variance).
-
-We emphasize that this does not contradict the well-known fact that kernel smoothing is an optimal method for nonparametric regression over e.g.
-H\"{o}lder balls.
-It simply reflects that in the standard nonparametric regression setup---which we adopt in the main part of this paper, and in which $P$ is assumed to be equivalent to Lebesgue measure---the biases of Laplacian eigenmaps and kernel smoothing are equivalent.
-On the other hand, when $f_0$ and $p$ satisfy some special relationship, such as the cluster assumption, the biases of these two methods can be quite different.
-There has been some work \textcolor{red}{(Rigollet, Wasserman, Niyogi, El Alaoui)} analyzing semi-supervised learning under various instantiations of the cluster assumption.
-However, a comprehensive analysis of various methods, including Laplacian eigenmaps and kernel smoothing, in this setting remains outstanding.
-
-\paragraph{Laplacian smoothing.}
-As mentioned previously, graph Laplacian smoothing uses the graph Laplacian $L_{n,\varepsilon}$ to form the penalty in a penalized least squares estimator.
-It can be calculated by one solve of a sparse, diagonally dominant linear system.
-Thus it should be much faster to compute than Laplacian eigenmaps, in which one must find (many) eigenvectors of $L_{n,\varepsilon}$.
-On the other hand, Laplacian smoothing is known to be minimax rate-optimal only in very limited regimes (over the Sobolev spaces $H^1(\mc{X})$ for $1 \leq d < 4$.) In contrast, we have shown that Laplacian eigenmaps is minimax rate-optimal for all $d$ (when $s = 1$) .
-We have also shown that Laplacian eigenmaps can adapt to higher-order smoothness, i.e.
-it can be minimax rate-optimal when $s > 1$.
-Thus, the known statistical properties of Laplacian eigenmaps are much stronger than those of Laplacian smoothing.
-
-\subsection{Future Work}
-We view our work can be viewed as a contribution both to the fields of nonparametric regression with series estimators, and to graph-based learning.
-We end our discussion by mentioning some open work in each of these directions.
-
-% (SB via AG): Add references, and clarify what is meant by used to.
-Much is known about classical spectral projection methods beyond their rate optimality.
-For instance: such estimators and tests exhibit \emph{sharp optimality}, meaning their risk is within a $(1 + o(1))$ factor of the optimal risk; they can adapt to unknown smoothness of the regression function; they can be used to estimate smooth functionals of the regression function;  finally, they can be used to form confidence sets in $L^2(P)$.
-It would be interesting to see if Laplacian eigenmaps could replicate the performance of classical methods in any, or all, of these problems.
-
-On the other hand, there are many variants of Laplacian eigenmaps worth considering.
-For instance, one can change the graph under consideration (e.g.
-by using the k-nearest neighbors), or the normalization of the graph Laplacian $L_{n,\varepsilon}$ (e.g.
-by using the symmetric normalized Laplacian).
-The former is practically useful, because it typically leads to connected graphs while always ensuring a given level of edge sparsity.
-In the latter, the graph Laplacian converges to a different limiting operator, which possesses different eigenvectors than $\Delta_P$ and thereby induces a different bias.
-We believe that under the setup we consider here, both methods will continue to be optimal.
+\input{sections/discussion}
 

### 2021-08-29 manuscripts/laplacian_eigenmaps/sections/discussion.tex
Partitioned off discussion section.

\section{Discussion}
\label{sec:discussion}

\subsection{Comparison with other estimators}
In this paper, we have motivated Laplacian eigenmaps by viewing it as a noisy approximation of a classical spectral projection method, which is its most obvious counterpart. We have shown that Laplacian eigenmaps inherits the optimality properties of its more classical counterpart. We now discuss the relationship between Laplacian eigenmaps and three other approaches to nonparametric regression. The first two---nonparametric least squares and kernel smoothing---are classical, whereas the third---Laplacian smoothing---makes use of the graph Laplacian in a different way than Laplacian eigenmaps.

% (SB via AG): Change the name of this to ``spectral projection'' or something like that.
\paragraph{Nonparametric least squares.}
The standard recommended alternative to spectral projection methods, when the distribution $P$ is considered non-uniform or unknown, is to do least-squares. For example, suppose instead of knowing $\psi_1,\psi_2,\ldots$, we had access only to eigenfunctions $\phi_1,\phi_2,\ldots$ of an unweighted Laplace-Beltrami operator $\Delta$. Then letting $\Phi_K = \mathrm{span}\{\phi_1,\ldots,\phi_K\}$, the least-squares estimator and test statistic
\begin{equation*}
\wt{f}_{\mathrm{LS}} := \argmin_{f \in \Phi_K} \|Y - f\|_n^2,\textrm{and}~~\wt{T}_{\mathrm{LS}} = \|\wt{f}\|_{\nu}^2
\end{equation*}
are still rate-optimal over $H_0^s(\mc{X})$. This holds true for both Model~\ref{def:model_flat_euclidean} and~\ref{def:model_manifold}, and is a consequence of our assumption that $p$ is bounded away from $0$.

However, this is not a totally satisfactory fix. For one thing, the least squares approach just outlined requires that we know the domain $\mc{X}$, in the strong sense that we know the Laplace-Beltrami operator $\Delta$ defined on $\mc{X}$. Domain knowledge is generally precious information, and such strong knowledge of $\mc{X}$ seems particularly unrealistic in the case where $\mc{X}$ is a manifold, and $\Delta$ is the manifold Laplace-Beltrami operator. Additionally, even if we know $\mc{X}$, diagonalizing the Laplace-Beltrami operator $\Delta$ is quite difficult for all but a few special domains, such as the unit cube $\mc{X} = [0,1]^d$ or torus $\mc{X} = \mathbb{T}^d$.

% (SB via AG): Maybe also worth discussing the ill-posed counterpart?

\paragraph{Kernel smoothing.}
It is also natural to ask whether the two stage estimator $T_{n,h}\wh{f}$ defined in Section~\ref{sec:out_of_sample} has any advantage over the simpler approach of directly kernel smoothing the responses, i.e. using the estimator $T_{n,h}Y$ (possibly for a different choice of $h$). In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{n,h}\wh{f}\|_P^2$ is of a strictly lower order than $\inf_{h'} \Ebb\|f_0 - T_{h',n}Y\|_P^2$. This is possible because Laplacian eigenmaps induces a completely different bias than kernel smoothing. For example, when $f_0$ and $p$ satisfy the so-called \emph{cluster} assumption--- e.g. $f_0$ is piecewise constant in high-density regions (clusters) of $p$--- then the bias of Laplacian eigenmaps can much smaller than that of kernel smoothing (for equivalent levels of variance).

We emphasize that this does not contradict the well-known fact that kernel smoothing is an optimal method for nonparametric regression over e.g. H\"{o}lder balls. It simply reflects that in the standard nonparametric regression setup---which we adopt in the main part of this paper, and in which $P$ is assumed to be equivalent to Lebesgue measure---the biases of Laplacian eigenmaps and kernel smoothing are equivalent. On the other hand, when $f_0$ and $p$ satisfy some special relationship, such as the cluster assumption, the biases of these two methods can be quite different. There has been some work \textcolor{red}{(Rigollet, Wasserman, Niyogi, El Alaoui)} analyzing semi-supervised learning under various instantiations of the cluster assumption. However, a comprehensive analysis of various methods, including Laplacian eigenmaps and kernel smoothing, in this setting remains outstanding.

\paragraph{Laplacian smoothing.}
As mentioned previously, graph Laplacian smoothing uses the graph Laplacian $L_{n,\varepsilon}$ to form the penalty in a penalized least squares estimator. It can be calculated by one solve of a sparse, diagonally dominant linear system. Thus it should be much faster to compute than Laplacian eigenmaps, in which one must find (many) eigenvectors of $L_{n,\varepsilon}$. On the other hand, Laplacian smoothing is known to be minimax rate-optimal only in very limited regimes (over the Sobolev spaces $H^1(\mc{X})$ for $1 \leq d < 4$.) In contrast, we have shown that Laplacian eigenmaps is minimax rate-optimal for all $d$ (when $s = 1$) . We have also shown that Laplacian eigenmaps can adapt to higher-order smoothness, i.e. it can be minimax rate-optimal when $s > 1$. Thus, the known statistical properties of Laplacian eigenmaps are much stronger than those of Laplacian smoothing.

\subsection{Future Work}
We view our work can be viewed as a contribution both to the fields of nonparametric regression with series estimators, and to graph-based learning. We end our discussion by mentioning some open work in each of these directions.

% (SB via AG): Add references, and clarify what is meant by used to.
Much is known about classical spectral projection methods beyond their rate optimality. For instance: such estimators and tests exhibit \emph{sharp optimality}, meaning their risk is within a $(1 + o(1))$ factor of the optimal risk; they can adapt to unknown smoothness of the regression function; they can be used to estimate smooth functionals of the regression function;  finally, they can be used to form confidence sets in $L^2(P)$. It would be interesting to see if Laplacian eigenmaps could replicate the performance of classical methods in any, or all, of these problems.

On the other hand, there are many variants of Laplacian eigenmaps worth considering. For instance, one can change the graph under consideration (e.g. by using the k-nearest neighbors), or the normalization of the graph Laplacian $L_{n,\varepsilon}$ (e.g. by using the symmetric normalized Laplacian). The former is practically useful, because it typically leads to connected graphs while always ensuring a given level of edge sparsity. In the latter, the graph Laplacian converges to a different limiting operator, which possesses different eigenvectors than $\Delta_P$ and thereby induces a different bias. We believe that under the setup we consider here, both methods will continue to be optimal.

### 2021-08-29 manuscripts/laplacian_eigenmaps/appendix.tex
Partitioned off discussion section.
Updated out-of-sample and manifold sections.

@@
 
-\begin{lemma}[Gagliardo-Nirenberg inequality]
+\begin{lemma}[Gagliardo-Nirenberg interpolation inequality]
 	\label{lem:gagliardo_nirenberg}
@@
 \end{lemma}
-\textcolor{red}{(TODO): Find the exact statement of this Lemma in \cite{leoni2017}.}
 
@@
 
-\textcolor{red}{(TODO): Check with Ryan and Siva that (i) they believe these facts, and (ii) these facts do not require additional justification.}
-
-\subsection{Upper bound on $L^2(P_n)$ risk}
-\label{subsec:upper_bound_l2pn_risk}
-\textcolor{red}{(TODO)}
+\section{Computational considerations}
+\label{sec:computational_considerations}
+Recall that when $s = 1$, we have shown that Laplacian eigenmaps is optimal when $\varepsilon \asymp (\log n/n)^{1/d}$ is (up to a constant) as small as possible while still ensuring the graph $G$ is connected.
+On the other hand, when $s > 1$, we can show Laplacian eigenmaps is optimal only when $\varepsilon = \omega(n^{-c})$ for some $c < 1/d$.
+For such a choice of $\varepsilon$, the average degree in $G$ will grow polynomially in $n$ as $n \to \infty$, and computing eigenvectors of the Laplacian of a graph will be more computationally intensive than if the graph were sparse \textcolor{red}{(reference)}.
+Thus Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and~\ref{thm:laplacian_eigenmaps_estimation_ho} can be seen as revealing a tradeoff between statistical and computational efficiency; although to be clear, we have no theoretical evidence that Laplacian eigenmaps \emph{fails} to adapt to higher-order smoothness when $\varepsilon \asymp (\log n/n)^{1/d}$---we simply cannot prove that it succeeds.
+\textcolor{red}{(TODO): If we decide it is worth our time to investigate the optimal choice of graph radius empirically, we can add a line here mentioning this.}
+
+Suppose one does choose $\varepsilon$ meaningfully larger than the connectivity threshold, as our theory requires when $s > 1$.
+We now discuss a procedure to efficiently compute an approximation to the Laplacian eigenmaps estimate, without changing the rate of convergence of the resulting estimator: \emph{edge sparsification}.
+By now there exist various methods see (e.g., the seminal papers of \citet{spielman2011,spielman2013,spielman2014}, or the overview by \citet{vishnoi2012} and references therein) to efficiently remove many edges from the graph $G$ while only slightly perturbing the spectrum of the Laplacian.
+Specifically such algorithms take as input a parameter $\sigma \geq 1$, and return a sparser graph $\wt{G}$, $E(\wt{G}) \subseteq E(G)$, with a Laplacian $\wt{L}_{n,\varepsilon}$ satisfying
+\begin{equation*}
+\frac{1}{\sigma} \cdot u^{\top} \wt{L}_{n,\varepsilon} u \leq u^{\top} L_{n,\varepsilon} u \leq \sigma \cdot u^{\top} \wt{L}_{n,\varepsilon}u \quad \textrm{for all $u \in \Reals^n$.}
+\end{equation*}
+Let $\wt{f}$ be the Laplacian eigenmaps estimator computed using the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ .
+Because $\wt{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wt{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$, and consequently much faster to compute $\wt{f}$ than $\wh{f}$ \textcolor{red}{(reference needed)}.
+Statistically speaking, letting $\wt{\lambda}_k$ be the $k$th eigenvalue of $\wc{L}_{n,\varepsilon}$, we have that conditional on $X_1,\ldots,X_n$,
+\begin{equation*}
+\|\wt{f} - f_0\|_n^2 \leq \frac{\dotp{\wt{L}_{n,\varepsilon}^s f_0}{f_0}_n}{\wt{\lambda}_{K + 1}^s} + \frac{5K}{n} \leq \sigma^{2s} \frac{\dotp{\wt{L}_{n,\varepsilon}^s f_0}{f_0}_n}{\wt{\lambda}_{K + 1}^s} + \frac{5K}{n},
+\end{equation*}
+with probability at least $1 - \exp(-K)$.
+Consequently $\wt{f}$ has $L^2(P_n)$-error of at most $\sigma^{2s}$ times our upper bound on the error of $\wh{f}$, and for any choice of $\sigma$ that is constant in $n$ the estimator $\wt{f}$ will also be rate-optimal.
+
+In fact the aforementioned edge sparsification algorithms are overkill for our needs.
+For one thing, they are designed to work when $\sigma$ is very close to $1$, whereas in order for $\wc{f}$ to be rate-optimal setting $\sigma$ to be any constant greater than $1$, say $\sigma = 2$, is sufficient.
+Additionally, edge sparsification algorithms are traditionally designed to work in the worst-case, where no assumptions are made on the structure of the graph $G$.
+But the geometric graphs we consider in this paper exhibit a special structure, in which very roughly speaking no single edge is a bottleneck.
+As pointed out by~\citet{sadhanala16b}, in this special case there are far simpler and faster methods for sparsification, which at least empirically seem to do the job.

### 2021-08-29 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Updated out-of-sample and manifold sections.

@@
 	\item \emph{Sub-critical Sobolev spaces}.
-We do not require that the regularity of the Sobolev space satisfy $s > d/2$, a condition often seen in the literature.
+Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and~\ref{thm:laplacian_eigenmaps_estimation_ho} do not require that the smoothness index $s$ of the Sobolev space satisfy $s > d/2$, a condition often seen in the literature.
 In the sub-critical regime $s \leq d/2$, the Sobolev space $H^s(\mc{X})$ is quite irregular.
@@
 Likewise, certain estimators are ``off the table'', most notably RKHS-based methods such as thin-plate splines of degree $k \leq d/2$.
-Nevertheless, for random design regression with error measured in $\Leb^2(P)$-norm, the spectral projection estimator $\wt{f}$ obtains the standard minimax rates $n^{-2s/(2s + d)}$ for all values of $s$ and $d$.
-Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and \ref{thm:laplacian_eigenmaps_estimation_ho} show that the same is true with respect to Laplacian eigenmaps, with error measured in $\Leb^2(P_n)$-norm.
+Nevertheless, for random design regression with error measured in squared $\Leb^2(P)$-norm, the spectral projection estimator $\wt{f}$ obtains the standard minimax rates $n^{-2s/(2s + d)}$ for all values of $s$ and $d$.
+Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and \ref{thm:laplacian_eigenmaps_estimation_ho} show that the same is true with respect to Laplacian eigenmaps, when error is measured in squared $\Leb^2(P_n)$-norm.
 	\item \emph{Smoothness of design density}.
@@
 The requirement $p \in C^{s - 1}(\mc{X})$ is essential to showing that $\wh{f}$ enjoys the faster minimax rates of convergence when $s > 1$,  as we discuss in Section~\ref{subsec:analysis}.
+	\item \emph{Computational considerations}.
+The lower bound on $\varepsilon$ in~\ref{asmp:parameters_estimation_ho} will result in a dense neighborhood graph $G$, meaning the average degree of $G$ will grow polynomially in the sample size $n$ as $n \to \infty$.
+As compared to a sparse $G$, this results in more non-zero entries in the graph Laplacian, and increases the computational burden involved in computing $\wh{f}$.
+To address this issue, in Appendix~\ref{sec:computational_considerations} we review some approaches to \emph{spectral sparsification}, in which one efficiently computes a sparse graph $\wc{G}$ that approximates $G$ in a spectral sense (see).
+Letting $\wc{f}$ be the Laplacian eigenmaps estimator computed with respect to the sparsified graph $\wc{G}$, we provide upper bounds on $\|\wc{f} - f_0\|_n^2$, which show that under mild conditions on $\wc{G}$---which provably achieved by many techniques for spectral sparsification---the estimator $\wc{f}$ achieves the same rates of convergence as $\wh{f}$.
 \end{itemize}
@@
 
-% (AG 8/27/21): Following chunk pasted here temporarily.
-We conclude this section by pointing out that it is somewhat remarkable how Laplacian eigenmaps can take advantage of higher-order smoothness in $f_0$.
-The sharpest known results \citep{calder2019,cheng2021} show that the squared error $\|v_k - \psi_k\|_n^2$ converges to $0$ at a rate of $n^{-2/(4 + d)}$).
-Naively applying these results, one can show that $\wh{f}$ to $\wt{f}$, but only at a rate far slower than the optimal rates for regression.
-Of course when the number of eigenvectors $K$ increases with $n$, as is necessary to optimally balance bias and variance of Laplacian eigenmaps, the issue only gets worse.
-Clearly, as a method for estimation and testing, the rate of convergence of Laplacian eigenmaps is much better than the rate implied by (what is currently known about) the concentration of individual eigenvectors around their continuum limits.
-
-\subsection{Computational considerations}
-Recall that when $s = 1$, we have shown that Laplacian eigenmaps is optimal when $\varepsilon \asymp (\log n/n)^{1/d}$ is (up to a constant) as small as possible while still ensuring the graph $G$ is connected.
-On the other hand, when $s > 1$, we can show Laplacian eigenmaps is optimal only when $\varepsilon = \omega(n^{-c})$ for some $c < 1/d$.
-For such a choice of $\varepsilon$, the average degree in $G$ will grow polynomially in $n$ as $n \to \infty$, and computing eigenvectors of the Laplacian of a graph will be more computationally intensive than if the graph were sparse \textcolor{red}{(reference)}.
-Thus Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and~\ref{thm:laplacian_eigenmaps_estimation_ho} can be seen as revealing a tradeoff between statistical and computational efficiency; although to be clear, we have no theoretical evidence that Laplacian eigenmaps \emph{fails} to adapt to higher-order smoothness when $\varepsilon \asymp (\log n/n)^{1/d}$---we simply cannot prove that it succeeds.
-\textcolor{red}{(TODO): If we decide it is worth our time to investigate the optimal choice of graph radius empirically, we can add a line here mentioning this.}
-
-Suppose one does choose $\varepsilon$ meaningfully larger than the connectivity threshold, as our theory requires when $s > 1$.
-We now discuss a procedure to efficiently compute an approximation to the Laplacian eigenmaps estimate, without changing the rate of convergence of the resulting estimator: \emph{edge sparsification}.
-By now there exist various methods see (e.g., the seminal papers of \citet{spielman2011,spielman2013,spielman2014}, or the overview by \citet{vishnoi2012} and references therein) to efficiently remove many edges from the graph $G$ while only slightly perturbing the spectrum of the Laplacian.
-Specifically such algorithms take as input a parameter $\sigma \geq 1$, and return a sparser graph $\wt{G}$, $E(\wt{G}) \subseteq E(G)$, with a Laplacian $\wt{L}_{n,\varepsilon}$ satisfying
-\begin{equation*}
-\frac{1}{\sigma} \cdot u^{\top} \wt{L}_{n,\varepsilon} u \leq u^{\top} L_{n,\varepsilon} u \leq \sigma \cdot u^{\top} \wt{L}_{n,\varepsilon}u \quad \textrm{for all $u \in \Reals^n$.}
-\end{equation*}
-Let $\wt{f}$ be the Laplacian eigenmaps estimator computed using the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ .
-Because $\wt{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wt{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$, and consequently much faster to compute $\wt{f}$ than $\wh{f}$ \textcolor{red}{(reference needed)}.
-Statistically speaking, letting $\wt{\lambda}_k$ be the $k$th eigenvalue of $\wc{L}_{n,\varepsilon}$, we have that conditional on $X_1,\ldots,X_n$,
-\begin{equation*}
-\|\wt{f} - f_0\|_n^2 \leq \frac{\dotp{\wt{L}_{n,\varepsilon}^s f_0}{f_0}_n}{\wt{\lambda}_{K + 1}^s} + \frac{5K}{n} \leq \sigma^{2s} \frac{\dotp{\wt{L}_{n,\varepsilon}^s f_0}{f_0}_n}{\wt{\lambda}_{K + 1}^s} + \frac{5K}{n},
-\end{equation*}
-with probability at least $1 - \exp(-K)$.
-Consequently $\wt{f}$ has $L^2(P_n)$-error of at most $\sigma^{2s}$ times our upper bound on the error of $\wh{f}$, and for any choice of $\sigma$ that is constant in $n$ the estimator $\wt{f}$ will also be rate-optimal.
-
-In fact the aforementioned edge sparsification algorithms are overkill for our needs.
-For one thing, they are designed to work when $\sigma$ is very close to $1$, whereas in order for $\wc{f}$ to be rate-optimal setting $\sigma$ to be any constant greater than $1$, say $\sigma = 2$, is sufficient.
-Additionally, edge sparsification algorithms are traditionally designed to work in the worst-case, where no assumptions are made on the structure of the graph $G$.
-But the geometric graphs we consider in this paper exhibit a special structure, in which very roughly speaking no single edge is a bottleneck.
-As pointed out by~\citet{sadhanala16b}, in this special case there are far simpler and faster methods for sparsification, which at least empirically seem to do the job.
+% (AG 8/29/21): Following chunk commented out but not deleted, in case it proves useful at a later point.
+% We conclude this section by pointing out that, at first glance, it may be somewhat remarkable that Laplacian eigenmaps can ever attain optimal rates of convergence.
+The sharpest known results \citep{calder2019,cheng2021} show that the squared error with which a graph Laplacian eigenvector approximates its continuum limit, $\|v_k - \psi_k\|_n^2$,  converges to $0$ at a rate of $C_k n^{-2/(4 + d)}$, for a constant $C_k \to \infty$ as $k \to \infty$.
+For any fixed $k$, this is slower than the optimal estimation or testing rates, and when the number of eigenvectors $K$ increases with $n$, as is necessary to optimally balance bias and variance of Laplacian eigenmaps, the issue only gets worse.
+Moreover, there are reasons to believe that this upper bound on $\|v_k - \psi_k\|_n^2$ is unimprovable (see~\cite{cheng2021}).
+If so, this suggests that Laplacian eigenmaps can learn regression function more efficiently, from a statistical perspective, than Laplacian eigenvectors can learn population-level eigenfunctions.
+Put more simply: regression is a forgiving problem.
 
@@
 In this section we consider the manifold setting, where $(X_1,Y_1),\ldots,(X_n,Y_n)$ are observed according to Model~\ref{def:model_manifold}.
-A theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
-We build on this work by showing that when $f_0 \in H^s(\mc{X})$ and $P$ is supported on a manifold, Laplacian eigenmaps achieve the sharper minimax estimation and testing rates reviewed in Section~\ref{subsec:minimax_rates_sobolev}.
+In this setting, it is known that the minimax rates depend only on the intrinsic dimension $m$; more specifically,~\cite{bickel2007,ariascastro2018} show that for H\"{o}lder classes, the minimax estimation rate is $n^{-2s/(2s + m)}$ and the testing rate is $n^{-4s/(4s + m)}$\footnote{Although~\cite{ariascastro2018} considers density testing, usual arguments regarding equivalence of experiments~\cite{brown96} imply that the same rates apply to regression testing}.
+On the other hand, a theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+We build on this work by showing that when $f_0 \in H^s(\mc{X})$ and $P$ is supported on a manifold, Laplacian eigenmaps achieve the sharper minimax estimation and testing rates.
 
@@
 Unlike in the flat-Euclidean case, since Model~\ref{def:model_manifold} assumes that $\mc{X}$ is boundaryless it is easy to deal with the first-order $(s = 1)$ and higher-order $(s > 1)$ cases all at once.
-A more important distinction between the results of this section and those of Section~\ref{sec:minimax_optimal_laplacian_eigenmaps} is that we will establish Laplacian eigenmaps is optimal only when the regression function $f_0 \in H^s(\mc{X};M)$ for $s \leq 3$.
+A more important distinction between the results of this section and those of Section~\ref{sec:minimax_optimal_laplacian_eigenmaps} is that we will establish Laplacian eigenmaps is optimal only when the regression function $f_0 \in H^s(\mc{X};M)$ for $s \in \{1,2,3\}$.
 Otherwise, this section will proceed in a similar fashion to Section~\ref{sec:higher_order_sobolev_classes}.
@@
 
-\begin{itemize}
-	\item The proofs of Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold} follow very similarly to the full-dimensional setting.
-The difference is that when $\mc{X}$ is a manifold with intrinsic dimension $m$, we can prove analogous results to Propositions~\ref{prop:graph_seminorm_fo}-\ref{prop:graph_eigenvalue}, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
-	\item Unlike in the full-dimensional case, in the manifold setting our upper bounds on the estimation and testing error of Laplacian eigenmaps match the minimax rate only when $s \leq 3$.
+Unlike in the full-dimensional case, in the manifold setting our upper bounds on the estimation and testing error of Laplacian eigenmaps match the minimax rate only when $s \leq 3$.
 When $s \geq 4$, the containment $H^s(\mc{X};1) \subset H^{3}(\mc{X};1)$ (taking the radius of the Sobolev ball $M = 1$ for simplicity) implies that the Laplacian eigenmaps estimator $\wh{f}$ has in-sample mean-squared error of at most $n^{-6/(6 + m)}$, and that the Laplacian eigenmaps test has small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-12/(12 + d)}$; however, these are slower than the usual minimax rates.
 
-	We now explain this discrepancy between our rates in the flat Euclidean and manifold settings.
-At a high level, thinking of the graph $G$ as an estimate of the manifold $\mc{X}$, we incur some error in this estimate by using Euclidean distance rather than geodesic distance to form the edges of $G$.
+We now explain this difference between the flat Euclidean and manifold settings.
+At a high level, thinking of the graph $G$ as an estimate of the manifold $\mc{X}$, we incur some error by using Euclidean distance rather than geodesic distance to form the edges of $G$.
 This is in contrast with the full-dimensional setting, where the Euclidean metric exactly coincides with the geodesic distance for all points $x,z \in \mc{X}$ that are sufficiently close to each other and far from the boundary of $\mc{X}$.
@@
 
-	\item Finally, as in the full-dimensional case, when the intrinsic dimension $m > 4$ we cannot choose the graph radius $\varepsilon$ and number of eigenvectors $K$ to optimally balance bias and variance.
+Finally, as in the full-dimensional case, when the intrinsic dimension $m > 4$ we cannot choose the graph radius $\varepsilon$ and number of eigenvectors $K$ to optimally balance bias and variance.
 Instead, reasoning as in the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} shows that when $1 \leq s \leq 3$, the Laplacian eigenmaps test has critical radius as given by~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_ho_suboptimal}, but with the ambient dimension $d$ replaced by $m$.
-\end{itemize}
 % (SB via AG): Maybe some comment on the difference between testing and estimation.
@@
 \label{sec:out_of_sample}
-Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity} show that $\wh{f}$ is a minimax optimal estimator over Sobolev spaces.
-However, as mentioned previously we have measured loss \emph{in-sample}---that is, measured in $\Leb^2(P_n)$ norm---whereas \emph{out-of-sample} error---error measured in $L^2(P)$ norm---is the more typical metric in the random design setup.
+Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity} show that $\wh{f}$ is a minimax optimal estimator over Sobolev spaces, when the loss is  \emph{in-sample}---that is, measured in $\Leb^2(P_n)$ norm.
+In this section we consider \emph{out-of-sample} error---error measured in $L^2(P)$ norm---which is the more typical metric used in the study of nonparametric regression with random design.
 
-Of course, the Laplacian eigenmaps estimator is only defined at the observed design points $X_1,\ldots,X_n$, and to measure its error in $L^2(P)$ norm we must first extend it to be defined over all of $\Xset$.
+Of course, the Laplacian eigenmaps estimator is only defined at the design points $\{X_1,\ldots,X_n\}$, and to measure error in $L^2(P)$ norm we must first extend it to be defined over the entire domain $\Xset$.
 We propose a simple method, kernel smoothing, to do the job.
@@
 \end{equation*}
-where $d_{Q,h}(x) := \int_{\Omega} \psi\bigl(\|z - x\|/\varepsilon\bigr) \,dQ(z)$.
+where $d_{Q,h}(x) := \int_{\Omega} \psi\bigl(\|z - x\|/h\bigr) \,dQ(z)$.
 For convenience, we will write $T_{n,h}f(x) := T_{P_n,h}f(x)$, and $d_{n,h}(x) := n \cdot d_{P_n,h}(x)$.
 We extend the Laplacian eigenmaps estimator by passing the kernel smoother $T_{n,h}$ over it, that is we consider the estimator $T_{n,h}\wh{f}$, which is defined at every $x \in \mc{X}$ (indeed, at every $x \in \Rd$).
-Note that ``extension'' here is a slight abuse of nomenclature, since $T_{n,h}\wh{f}(X_i)$ and $\wh{f}_i$ may not agree in-sample.
+Note that ``extension'' here is a slight abuse of nomenclature, since $T_{n,h}\wh{f}$ and $\wh{f}$ may not agree in-sample.
 
@@
 	\label{lem:kernel_smoothing_insample}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $\wc{f} \in L^2(P_n)$, $f_0 \in H^1(\mc{X})$ and $p \in C^1(\mc{X})$.
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f_0 \in H^1(\mc{X};M)$ and $p \in C^1(\mc{X})$.
 If the kernel smoothing estimator $T_{n,h}\wc{f}$ is computed with a kernel $\psi$ satisfying~\ref{asmp:kernel} and bandwidth $h$ satisfying~\ref{asmp:bandwidth}, it holds that
@@
 Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data, by a factor of $h^2$.
-On the other hand the bias term is typical.
-When $\psi$ is an order-$s$ kernel, a standard analysis shows that the $\|T_{P,h}f_0 - f_0\|_P^2 \lesssim \varepsilon^{2s}$.
+On the other hand the bias term is typical, and when $\psi$ is an order-$s$ kernel a standard analysis (see Lemma~\ref{lem:kernel_smoothing_bias}) shows that $\|T_{P,h}f_0 - f_0\|_P^2 \lesssim |f|_{H^s(\mc{X})}^2 h^{2s}$.
 \begin{enumerate}[label=(K\arabic*)]
@@
 	\label{asmp:ho_kernel}
-	The kernel function $\psi$ is an order-$s$ kernel, meaning that it satisfies
+	The kernel function $\psi$ is an order-$s$ kernel, meaning it satisfies
 	\begin{equation*}
@@
 
-Choosing $h \asymp n^{-1/(2(s - 1) + d)}$ balances the kernel smoothing bias and variance terms in~\eqref{eqn:kernel_smoothing_insample}, and implies that
+Choosing $h \asymp n^{-1/(2(s - 1) + d)}$ balances the kernel smoothing bias and variance terms in~\eqref{eqn:kernel_smoothing_insample}, and implies that for $f_0 \in H^s(\mc{X};M)$,
 \begin{equation}
 \label{eqn:kernel_smoothing_insample2}
-\|T_{n,h}\wc{f} - f_0\|_P^2 \leq C\biggl(\|\wc{f} - f_0\|_n^2 + \frac{1}{\delta}n^{-2s/(2(s - 1) + d)}\biggr).
+\|T_{n,h}\wc{f} - f_0\|_P^2 \leq C\biggl(\|\wc{f} - f_0\|_n^2 + \frac{M^2}{\delta}n^{-2s/(2(s - 1) + d)}\biggr).
 \end{equation}
 This analysis tells us that the additional error incurred by passing a kernel smoother over an in-sample estimator $\wc{f}$ is negligible compared to the minimax rate of estimation.
-Consequently, if $\wc{f}$ converges at the minimax rate in in-sample mean squared error, then $T_{n,h}\wc{f}$ will converge at the minimax rate in $L^2(P)$.
+Consequently, if $\wc{f}$ converges at the minimax rate in squared $L^2(P_n)$-norm, then $T_{n,h}\wc{f}$ will converge at the minimax rate in squared $L^2(P)$-norm.
 It follows immediately from Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} (when $s = 1$) or Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} (when $s > 1$), that $T_{n,h}\wh{f}$ achieves the optimal rate of convergence in $L^2(P)$.
@@
 		\begin{equation*}
-		\|T_{n,h}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2s/(2s + d)}.
+		\|T_{n,h}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2/(2 + d)}.
 		\end{equation*}
@@
 \begin{itemize}
-	\item Since $T_{n,h}\wh{f}$ is defined out-of-sample, we can use sample splitting or cross validation methods to tune hyperparameters, which we could not do for the original estimator $\wh{f}$.
+	\item \emph{Data-driven parameter tuning}.
+Since $T_{n,h}\wh{f}$ is defined out-of-sample, we can use sample splitting or cross validation methods to tune hyperparameters, which we could not do for the original estimator $\wh{f}$.
 For instance, we can (i) split the sample into two halves, (ii) use the first half to compute $T_{n,h}\wh{f}$ for various values of $\varepsilon$, $h$, and $K$, (iii) choose the optimal values of these three hyperparameters by minimizing error on the held out set.
 Practically speaking, cross-validation is one of the most common approaches to choosing hyperparameters.
-Theoretically, it is known \textcolor{red}{(references)} that choosing hyper-parameters through sample splitting can result in estimators that optimally adapt to the order of regularity $s$.
+Theoretically, it is known that choosing hyper-parameters through sample splitting can result in estimators that optimally adapt to the order of regularity $s$ \citep{gyorfi2006}.
 In other words, it leads to estimators that are rate-optimal (up to $\log n$ factors), even when $s$ is unknown.
-Similar arguments should imply that $T_{n,h}\wh{f}$ is adaptive in this sense when $\varepsilon,h$ and $K$ are chosen by sample splitting.
-\textcolor{red}{(TODO): Siva would like me to fill in some details?}
-	\item There exist other approaches to extending a function $f$ from its $\{f(X_1),\ldots,f(X_n)\}$: for instance, minimum-norm interpolation in some RKHS or Banach space \textcolor{blue}{(Rieger2008, Belkin2018)}, or 1-nearest neighbors regression \textcolor{blue}{(citation)}.
+We believe similar arguments should imply that $T_{n,h}\wh{f}$ is adaptive in this sense when $\varepsilon,h$ and $K$ are chosen by sample splitting, but the details should be worked out carefully.
+	\item \emph{Alternatives to extension by kernel smoothing}.
+There exist other approaches to extending a function $f$ from its evaluations $\{f(X_1),\ldots,f(X_n)\}$: for instance, minimum-norm interpolation in some RKHS or Banach space or 1-nearest neighbors regression.
 We consider extension by kernel smoothing because it is a simple and statistically optimal procedure that does not require any knowledge of the domain $\mc{X}$ or distribution $P$---as we have argued, this latter property is one of the main selling points of Laplacian eigenmaps as a tool for nonparametric regression.
 One potential drawback to kernel smoothing is that it can change the value of the estimate at data, meaning $\wh{f} \neq (\wc{f}(X_1),\ldots, \wc{f}(X_n))$.
-More sophisticated procedures such as Nystr\"{o}m extension of the eigenvectors $v_1,\ldots,v_K$ \textcolor{blue}{(citation)} inherit the generality of kernel smoothing while being genuine interpolators; however, their properties are substantially more difficult to analyze.
+More sophisticated procedures based on Nystr\"{o}m extension of the eigenvectors $v_1,\ldots,v_K$~\cite{fowlkes2004} would inherit the generality of kernel smoothing while being genuine interpolators; however, their properties are substantially more difficult to analyze.
 
@@
 In this section we empirically demonstrate that Laplacian Eigenmaps is a reasonably good alternative to spectral projection, even when $n$ is only moderately large.
-In order to compare the two methods, in our experiments we stick to simple settings where we can compute eigenfunctions of $\Delta_P$, and thus the spectral projection estimator.
-Of course in general, it is not easy to compute these eigenfunctions: hence the appeal of Laplacian Eigenmaps.
+In order to compare the two methods, in our experiments we stick to the simple case where $P$ is the uniform distribution over $\mc{X} = [0,1]^d$, and we have simple closed-form expressions for the eigenfunctions of $\Delta_P$.
+In general, it is not easy to analytically compute these eigenfunctions, which is part of the appeal of Laplacian Eigenmaps.
 
@@
 
-In our first experiment, we compare the mean-squared error of Laplacian eigenmaps to that of its classical spectral projection counterpart.
-We vary the sample size from $n = 1000$ to $n = 4000$; sample $n$ design points $X_1,\ldots,X_n$ from the uniform distribution on the cube $[-1,1]^d$; and sample responses $Y_i$ according to~\eqref{eqn:model} with regression function $f_0 = M/\lambda_K^{s/2} \cdot \psi_K$ for $K \asymp n^{d/(2s + d)}$; the pre-factor $M/\lambda_K^{s/2}$ is chosen so that $|f_0|_{H^s(\mc{X})}^2 = M^2$.
-In Figure~\ref{fig:fig1} we show the in-sample mean-squared error of Laplacian eigenmaps and a classical spectral projection estimator as a function of $n$, for different dimensions $d$ and order of smoothness $s$.
+\paragraph{Estimation.}
+In our first experiment, we compare the mean-squared error of the Laplacian eigenmaps estimator $\wh{f}$  to that of its population-level counterpart $\wt{f}$.
+We vary the sample size from $n = 1000$ to $n = 4000$; sample $n$ design points $X_1,\ldots,X_n$ from the uniform distribution on the cube $[-1,1]^d$; and sample responses $Y_i$ according to~\eqref{eqn:model} with regression function $f_0 = M/\lambda_K^{s/2} \cdot \psi_K$ for $K \asymp n^{d/(2s + d)}$ (the pre-factor $M/\lambda_K^{s/2}$ is chosen so that $|f_0|_{H^s(\mc{X})}^2 = M^2$).
+In Figure~\ref{fig:fig1} we show the in-sample mean-squared error of the two estimators as a function of $n$, for different dimensions $d$ and order of smoothness $s$.
 We see that all estimators have mean-squared error converging to zero at roughly the minimax rate.
-We also see that the mean-squared error of Laplacian Eigenmaps gets closer to that of spectral projection as $n$ gets larger.
-The fact that spectral projection outperform Laplacian Eigenmaps
+While unsurprisingly population-level spectral projection has smaller error, generally speaking that the error of Laplacian Eigenmaps gets closer to that of population-level spectral projection as $n$ gets larger.
 
-We also compare the error, over a held-out test set, of Laplacian eigenmaps plus kernel smoothing to the spectral projection estimator.
-The out-of-sample mean squared error of the two estimators is very similar to the in-sample mean-squared error.
-This supports our theoretical claim that the additional error incurred by kernel smoothing of Laplacian Eigenmaps is negligible.
+We also compare the out-of-sample estimation error, over a held-out test set, of Laplacian eigenmaps plus kernel smoothing to the spectral projection estimator.
+The story is much the same, supporting our theoretical claim that the additional error incurred by kernel smoothing of Laplacian Eigenmaps is negligible.
 
-In our second experiment, we compare tests using Laplacian eigenmaps and spectral projection test statistics.
+\paragraph{Testing.} 
+In our second experiment, we compare the Laplacian eigenmaps test $\varphi$ against the  population-level spectral projection test $\wt{\varphi}$.
 The setup, in terms of $n$ and $P$, is the same as that of our first experiment.
@@
 Then we take $\bar{\epsilon} = 1/100 \cdot \sum_{b = 1}^{100}\epsilon_b$ to be our empirical measure of worst-case risk.
-In Figure~\ref{fig:fig2}, we see that the critical radii of both Laplacian eigenmaps and spectral projection tests are quite close to each other, and converge to $0$ at roughly the minimax rate.
+In Figure~\ref{fig:fig2}, we see that the critical radii of both Laplacian eigenmaps and spectral projection tests are quite close to each other, and converge at roughly the minimax rate.
 \begin{figure*}[tb]
@@
 
-These experiments demonstrate that in terms of statistical error, Laplacian eigenmaps methods are reasonable replacements for spectral projection methods.
+\paragraph{Tuning parameters.}
+Our first two experiments demonstrate that in terms of statistical error, Laplacian eigenmaps methods are reasonable replacements for spectral projection methods.
 Laplacian eigenmaps depends on two tuning parameters, and in our final experiment we investigate the importance of both, focusing now on estimation.

### 2021-08-29 manuscripts/laplacian_eigenmaps/sections/discussion.tex
Updated out-of-sample and manifold sections.

@@
 
-\subsection{Comparison with other estimators}
-In this paper, we have motivated Laplacian eigenmaps by viewing it as a noisy approximation of a classical spectral projection method, which is its most obvious counterpart.
-We have shown that Laplacian eigenmaps inherits the optimality properties of its more classical counterpart.
-We now discuss the relationship between Laplacian eigenmaps and three other approaches to nonparametric regression.
-The first two---nonparametric least squares and kernel smoothing---are classical, whereas the third---Laplacian smoothing---makes use of the graph Laplacian in a different way than Laplacian eigenmaps.
+In this work, we have derived upper bounds on the rates of convergence for methods based on Laplacian Eigenmaps, which imply that in various settings these methods are minimax rate-optimal over Sobolev classes.
+Importantly, these upper bounds hold under nonparametric conditions on the design density $p$, and allow for $p$ to be unknown and, potentially, supported on a low-dimensional manifold.
+Our results help explain the practical success of methods which leverage graph Laplacian eigenvectors in downstream learning problems.
+They also distinguish Laplacian Eigenmaps from more traditional spectral projection procedures, which rely on a density-dependent basis and thus require the density be known a priori.
 
-% (SB via AG): Change the name of this to ``spectral projection'' or something like that.
-\paragraph{Nonparametric least squares.}
-The standard recommended alternative to spectral projection methods, when the distribution $P$ is considered non-uniform or unknown, is to do least-squares.
-For example, suppose instead of knowing $\psi_1,\psi_2,\ldots$, we had access only to eigenfunctions $\phi_1,\phi_2,\ldots$ of an unweighted Laplace-Beltrami operator $\Delta$.
-Then letting $\Phi_K = \mathrm{span}\{\phi_1,\ldots,\phi_K\}$, the least-squares estimator and test statistic
-\begin{equation*}
-\wt{f}_{\mathrm{LS}} := \argmin_{f \in \Phi_K} \|Y - f\|_n^2,\textrm{and}~~\wt{T}_{\mathrm{LS}} = \|\wt{f}\|_{\nu}^2
-\end{equation*}
-are still rate-optimal over $H_0^s(\mc{X})$.
-This holds true for both Model~\ref{def:model_flat_euclidean} and~\ref{def:model_manifold}, and is a consequence of our assumption that $p$ is bounded away from $0$.
+Of course, there do exist other methods for nonparametric regression which achieve optimal rates of convergence under similar (or indeed weaker) conditions on $p$.
+These include other graph-based approaches---graph Laplacian regularization---methods besides spectral projection---e.g.
+kernel smoothing, local polynomial regression, thin-plate splines---and continuum spectral projection methods which use the eigenfunctions of an operator defined independently of $p$.
+To be clear, we do not advocate Laplacian eigenmaps over these alternatives.
+Rather, we view our results as theoretically justifying a place for Laplacian eigenmaps in the nonparametric regression toolbox.
 
-However, this is not a totally satisfactory fix.
-For one thing, the least squares approach just outlined requires that we know the domain $\mc{X}$, in the strong sense that we know the Laplace-Beltrami operator $\Delta$ defined on $\mc{X}$.
-Domain knowledge is generally precious information, and such strong knowledge of $\mc{X}$ seems particularly unrealistic in the case where $\mc{X}$ is a manifold, and $\Delta$ is the manifold Laplace-Beltrami operator.
-Additionally, even if we know $\mc{X}$, diagonalizing the Laplace-Beltrami operator $\Delta$ is quite difficult for all but a few special domains, such as the unit cube $\mc{X} = [0,1]^d$ or torus $\mc{X} = \mathbb{T}^d$.
-
-% (SB via AG): Maybe also worth discussing the ill-posed counterpart?
-
-\paragraph{Kernel smoothing.}
-It is also natural to ask whether the two stage estimator $T_{n,h}\wh{f}$ defined in Section~\ref{sec:out_of_sample} has any advantage over the simpler approach of directly kernel smoothing the responses, i.e.
-using the estimator $T_{n,h}Y$ (possibly for a different choice of $h$).
-In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{n,h}\wh{f}\|_P^2$ is of a strictly lower order than $\inf_{h'} \Ebb\|f_0 - T_{h',n}Y\|_P^2$.
+That being said, Laplacian eigenmaps can have advantages over each of the aforementioned approaches. We now conclude by outlining some of these advantages, limiting our discussion to estimation:
+\begin{itemize}
+	\item \emph{Optimality over low-smoothness Sobolev spaces}.
+Graph Laplacian regularization achieves minimax optimal rates over $H^1(\mc{X})$ only when $d \in \{1,2,3,4\}$ \citep{sadhanala16, green2021}; when $d \geq 5$, graph Laplacian regularization appears to be suboptimal.
+In contrast, Laplacian eigenmaps is optimal for all dimensions $d$.
+An even more stark contrast can be drawn between Laplacian eigenmaps and the first order thin-plate spline, the solution of the variation problem
+	\begin{equation}
+	\underset{f \in H^1(\mc{X})}{\textrm{minimize}} \quad  \|{\bf Y} - f\|_n^2 + \lambda \|f\|_{H^1(\Reals^d)}^2,
+	\end{equation}
+	which is optimal and indeed well-posed only when $d = 1$.
+This appears to be related to a general phenomenon described by~\cite{dhillon2013,dicker2017}, in which principal components analysis followed by ordinary least-squares is shown to be always competitive with, and sometimes much better than, ridge regression.
+	\item \emph{Manifold adaptivity}.
+An oft-recommended alternative to the population-level spectral projection methods considered in Section~\ref{subsec:spectral_projection} is to run ordinary-least squares regression using eigenfunctions of a density-independent operator, such as the Laplacian operator $\Delta = \sum_{i = 1}^{d} \partial^2f/\partial x_i^2$.
+Under the conditions of Model~\ref{def:model_flat_euclidean}, such a method will indeed be rate-optimal, though the upper bounds may come with undesirably large constants if $p$ is very non-uniform.
+However, in contrast with Laplacian eigenmaps, under Model~\ref{def:model_manifold} this method cannot achieve the faster minimax rates of convergence.
+	\item \emph{Density adaptivity}.
+It is natural to ask whether the two stage estimator $T_{n,h}\wh{f}$ defined in Section~\ref{sec:out_of_sample} has any advantage over the simpler approach of directly kernel smoothing the responses, i.e.
+using the estimator $T_{n,h}{\bf Y}$ (possibly for a different choice of $h$).
+In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{n,h}\wh{f}\|_P^2$ is of a strictly lower order than $\inf_{h'} \Ebb\|f_0 - T_{h',n}{\bf Y}\|_P^2$.
 This is possible because Laplacian eigenmaps induces a completely different bias than kernel smoothing.
@@
 
-We emphasize that this does not contradict the well-known fact that kernel smoothing is an optimal method for nonparametric regression over e.g.
+	We emphasize that this does not contradict the well-known optimality properties of kernel smoothing over e.g.
 H\"{o}lder balls.
-It simply reflects that in the standard nonparametric regression setup---which we adopt in the main part of this paper, and in which $P$ is assumed to be equivalent to Lebesgue measure---the biases of Laplacian eigenmaps and kernel smoothing are equivalent.
-On the other hand, when $f_0$ and $p$ satisfy some special relationship, such as the cluster assumption, the biases of these two methods can be quite different.
-There has been some work \textcolor{red}{(Rigollet, Wasserman, Niyogi, El Alaoui)} analyzing semi-supervised learning under various instantiations of the cluster assumption.
-However, a comprehensive analysis of various methods, including Laplacian eigenmaps and kernel smoothing, in this setting remains outstanding.
-
-\paragraph{Laplacian smoothing.}
-As mentioned previously, graph Laplacian smoothing uses the graph Laplacian $L_{n,\varepsilon}$ to form the penalty in a penalized least squares estimator.
-It can be calculated by one solve of a sparse, diagonally dominant linear system.
-Thus it should be much faster to compute than Laplacian eigenmaps, in which one must find (many) eigenvectors of $L_{n,\varepsilon}$.
-On the other hand, Laplacian smoothing is known to be minimax rate-optimal only in very limited regimes (over the Sobolev spaces $H^1(\mc{X})$ for $1 \leq d < 4$.) In contrast, we have shown that Laplacian eigenmaps is minimax rate-optimal for all $d$ (when $s = 1$) .
-We have also shown that Laplacian eigenmaps can adapt to higher-order smoothness, i.e.
-it can be minimax rate-optimal when $s > 1$.
-Thus, the known statistical properties of Laplacian eigenmaps are much stronger than those of Laplacian smoothing.
-
-\subsection{Future Work}
-We view our work can be viewed as a contribution both to the fields of nonparametric regression with series estimators, and to graph-based learning.
-We end our discussion by mentioning some open work in each of these directions.
-
-% (SB via AG): Add references, and clarify what is meant by used to.
-Much is known about classical spectral projection methods beyond their rate optimality.
-For instance: such estimators and tests exhibit \emph{sharp optimality}, meaning their risk is within a $(1 + o(1))$ factor of the optimal risk; they can adapt to unknown smoothness of the regression function; they can be used to estimate smooth functionals of the regression function;  finally, they can be used to form confidence sets in $L^2(P)$.
-It would be interesting to see if Laplacian eigenmaps could replicate the performance of classical methods in any, or all, of these problems.
-
-On the other hand, there are many variants of Laplacian eigenmaps worth considering.
-For instance, one can change the graph under consideration (e.g.
-by using the k-nearest neighbors), or the normalization of the graph Laplacian $L_{n,\varepsilon}$ (e.g.
-by using the symmetric normalized Laplacian).
-The former is practically useful, because it typically leads to connected graphs while always ensuring a given level of edge sparsity.
-In the latter, the graph Laplacian converges to a different limiting operator, which possesses different eigenvectors than $\Delta_P$ and thereby induces a different bias.
-We believe that under the setup we consider here, both methods will continue to be optimal.
+Rather, in the standard nonparametric regression setup---which we adopt in the main part of this paper, and in which $P$ is assumed to be equivalent to Lebesgue measure---the biases of Laplacian eigenmaps and kernel smoothing happen to be equivalent.
+But in general this is not the case.
+\end{itemize}
+Grounding each of these three points on a firmer and more complete theoretical basis would be, in our view, a valuable direction for future work.

### 2021-08-29 manuscripts/laplacian_eigenmaps/sections/introduction.tex
Updated out-of-sample and manifold sections.

@@
 \cite{lee2016} consider the \emph{diffusion maps} estimator---which uses the eigenvectors of a differently normalized graph Laplacian---in a similar supervised setting to our own, but prove suboptimal rates of convergence.
-Finally, \citet{trillos2020,green2021} study a penalized least squares method for estimation, which uses a quadratic form involving the Laplacian as a penalty.
-This method is optimal only for the first-order Sobolev space $H^1(\mc{X})$, and even then only when $d \in \{1,2,3,4\}$, in contrast to the Laplacian eigenmaps estimator which is optimal for all $s$ and $d$; we discuss this contrast more in Section~\ref{sec:discussion}.
+
+Finally, \citet{trillos2020,green2021} study \emph{graph Laplacian regularization}, a graph-based penalized least squares method which uses a quadratic form involving the graph Laplacian as a penalty.
+Laplacian regularization can, like Laplacian eigenmaps, be cast as a spectral method.
+The difference is that Laplacian regularization induces some smooth shrinkage based on the eigenvalue, where Laplacian eigenmaps corresponds to a hard spectral cutoff.
+Interestingly, this difference leads to a stark contrast in the ultimate statistical properties of the two approaches.
+For instance, in~\cite{green2021} Laplacian regularization is shown to be optimal only for the first-order Sobolev space $H^1(\mc{X})$, and even then only when $d \in \{1,2,3,4\}$.
+In contrast, in this work we show that the Laplacian eigenmaps estimator is optimal for all $s$ and $d$.
 
@@
 In Section~\ref{sec:experiments} we examine the empirical behavior of Laplacian eigenmaps, and show that even at moderate sample sizes Laplacian eigenmaps is competitive with classical spectral projection.
-We conclude with some discussion in Section~\ref{sec:discussion}, comparing Laplacian eigenmaps to a few selected estimators and commenting on some computational considerations.
+We conclude with some discussion in Section~\ref{sec:discussion}.
 

### 2021-08-29 manuscripts/laplacian_eigenmaps/sections/preliminaries.tex
Updated out-of-sample and manifold sections.

@@
 
-On the other hand, the eigenvectors and eigenfunctions of $\Delta_P$ can also be used to give a spectral definition of Sobolev spaces. Consider the ellipsoid
+On the other hand, the eigenvectors and eigenfunctions of $\Delta_P$ can also be used to give a spectral definition of Sobolev spaces on general domains $\mc{X}$. Consider the ellipsoid
 \begin{equation}
@@
 To this end, we now give a pair of results, Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, which establish that classical spectral projection methods are statistically optimal for nonparametric estimation and testing over Sobolev ellipsoids.
-These results are generalizations of previously known upper bounds on the estimation and testing error of classical spectral methods~\citep{tsybakov08,ingster2009}, in that they hold under less stringent (nonparametric) conditions on the design distribution $P$.
+These results are generalizations of previously known upper bounds on the estimation and testing error of classical spectral methods~\citep{tsybakov08,ingster2009}, but they hold under less stringent (nonparametric) conditions on the design distribution $P$.
 This latter point is important, since as we have argued one attractive quality of Laplacian Eigenmaps is its adaptivity to general design distributions.

### 2021-08-31 manuscripts/laplacian_eigenmaps/appendix.tex
Made many small changes to Laplacian Eigenmaps paper. Also converted format back to something Siva would prefer.

--- 2021-08-23 manuscripts/laplacian_eigenmaps/appendix.tex
@@
 
-\subsection{Upper bound on Estimation Error of Laplacian Eigenmaps}
+\subsection{Upper bound on Estimation Error of Laplacian eigenmaps}
 
@@
 The bound holds for all $x \in \mc{X}_{k\varepsilon}$, and $f \in H^{s}(\mc{X})$, as long as $s \geq 2k + 1$.
-\textcolor{red}{(TODO): Make sure that $\sigma_{\eta}$ is defined somewhere.}
 \begin{lemma}
@@
 \end{lemma}
-We remark that when $k = 1$ and $f \in C^3(\mc{X})$ or $C^4(\mc{X})$, statements of this kind are well known \textcolor{red}{(references)}, and indeed stronger results---with $L^{\infty}(\mc{X})$ norm replacing $L^2(\mc{X})$ norm---hold.
-When dealing with the iterated Laplacian, and functions $f$ which are regular only in the Sobolev sense, the proof is somewhat more lengthy, but the spirit of the result is largely the same.
+We remark that when $k = 1$ and $f \in C^3(\mc{X})$ or $C^4(\mc{X})$, statements of this kind are well known, and indeed stronger results---with $L^{\infty}(\mc{X})$ norm replacing $L^2(\mc{X})$ norm---hold.
+When dealing with the iterated Laplacian, and functions $f$ which are regular only in the Sobolev sense, the proof is somewhat more lengthy, but in result is similar in spirit.
 
@@
 \end{equation}
-Finally, we recall a result of \textcolor{blue}{(Green 21)}: there exist constants $c_0$ and $C_0$ which do not depend on $M$, such that for all $\varepsilon < c_0$ and for any $f \in H^1(\mc{X};M)$,
+Finally, we recall a result of~\cite{green2021}: there exist constants $c_0$ and $C_0$ which do not depend on $M$, such that for all $\varepsilon < c_0$ and for any $f \in H^1(\mc{X};M)$,
 \begin{equation}
@@
 and upper bound each energy separately.
-For the first term, we add and substract $\sigma_{\eta}^k\Delta_P^kf(x)$ and $\sigma_{\eta}^k\Delta_P^kf(x')$ within the integrand, then use the triangle inequality and \textcolor{red}{the symmetry between $x$ and $x'$} to deduce that
+For the first term, we add and substract $\sigma_{\eta}^k\Delta_P^kf(x)$ and $\sigma_{\eta}^k\Delta_P^kf(x')$ within the integrand, then use the triangle inequality and the symmetry between $x$ and $x'$ to deduce that
 \begin{equation}
@@
 \label{subsec:integrals}
-
-\textcolor{red}{(TODO): The language ``...for a Borel set $U \subseteq \mc{X}$...'' should be checked.}
 
@@
 \end{align}
-with the second inequality following from the proof of Lemma~1 in \textcolor{blue}{(Green 2021).} Finally, we notice that $|i\bj\setminus \{i \cup j\}| + 1 = k$, so that~\eqref{pf:graph_seminorm_bias2_1} gives the desired result.
+with the second inequality following from the proof of Lemma~1 in~\cite{green2021}.
+Finally, we notice that $|i\bj\setminus \{i \cup j\}| + 1 = k$, so that~\eqref{pf:graph_seminorm_bias2_1} gives the desired result.
 
@@
 This is analogous to the situation in the flat Euclidean model.
+
+In the proof of~\eqref{pf:graph_seminorm_manifold_1}-\eqref{eqn:manifold_nonlocal_seminorm}, we recall the following estimates from differential geometry: (i) letting $K_0$ be an upper bound on the absolute value of the sectional curvatures of $\mc{X}$, $K_0 \leq 2R$, and  letting (ii) $i_0$ be a lower bound on the injectivity radius of $\mc{X}$, $i_0 \geq \pi R$; see Proposition~1 of \cite{aamari2019}.
+Additionally, recall that for all $\delta < i_0$, the exponential map $\exp_x: B_m(0,\delta) \subset T_x(\mc{X}) \to B_{\mc{X}}(x,\delta) \subset \mc{X}$ is a diffeomorphism for all $x \in \mc{X}$.
+We shall therefore always assume $\varepsilon < i_0$.
 
@@
 Thus we have upper bounded $|L_{P,\varepsilon}f(x) - \wt{L}_{P,\varepsilon}f(x)|^2$ by the product of two terms, each of which we now suitably bound.
-
-To do so, we will use the following estimates, from Proposition 4 of \cite{trillos2019}: letting $R$ denote the reach of $\mc{X}$, for all $\|x' - x\| \leq R/2$,
+To do so, we will use the following estimate, from Proposition 4 of \cite{trillos2019}: for all $\|x' - x\| \leq R/2$,
 \begin{equation}
@@
 \end{equation}
+From here forward we will assume $\varepsilon < R/2$. 
 \paragraph{Upper bound on $A_1(x)$.}
-From here forward we will assume $\varepsilon < R/2$.
 Consequently $\eta(\|x' - x\|/\varepsilon) \geq \eta(d_{\mc{X}}(x',x)/\varepsilon)$.
@@
 \begin{equation*}
-\int_{\mc{X}} A_1(x) \,d\mu(x) \leq \frac{8L_{\eta}\nu_m\varepsilon^2}{R^2(m + 2)} \Bigl(1 + CmKR^2\Bigr) \varepsilon^{m + 2} |f|_{H^1(\mc{X})}^2 + C\|\eta\|_{\infty}\varepsilon^{m + 4} |f|_{H^1(\mc{X})}^2 =: C_5 \varepsilon^{m + 4}|f|_{H^1(\mc{X})}.
+\int_{\mc{X}} A_1(x) \,d\mu(x) \leq \frac{8L_{\eta}\nu_m\varepsilon^2}{R^2(m + 2)} \Bigl(1 + CmK_0R^2\Bigr) \varepsilon^{m + 2} |f|_{H^1(\mc{X})}^2 + C\|\eta\|_{\infty}\varepsilon^{m + 4} |f|_{H^1(\mc{X})}^2 =: C_5 \varepsilon^{m + 4}|f|_{H^1(\mc{X})}.
 \end{equation*}
@@
 \begin{equation*}
-\bigl|\mu(B_\mc{X}(x,\varepsilon))  - \omega_m \varepsilon^m\bigr|  \leq CmK\varepsilon^{m + 2},
-\end{equation*}
-where $K$ is an upper bound on the \textcolor{red}{sectional curvature} of $\mc{X}$. Plugging this back into~\eqref{pf:error_euclidean_distance_1}, we conclude that
-\begin{align*}
-\int_{\mc{X}} \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') & \leq  \frac{8L_{\eta}\varepsilon^2}{R^2}\Bigl[\omega_m\varepsilon^{m} + CmK\varepsilon^{m + 2}\Bigr] + \|\eta\|_{\infty} \Bigl[\omega_m(\wt{\varepsilon}^m - {\varepsilon}^m) + 2CmK\varepsilon^{m + 2}\Bigr] \\
-& \leq \frac{8L_{\eta}\varepsilon^2}{R^2}\Bigl[\omega_m\varepsilon^{m} + R^2CmK\varepsilon^{m}\Bigr] + \|\eta\|_{\infty} \varepsilon^{m + 2}\Bigl[\frac{27\omega_m}{R^2} + 2CmK\Bigr] \\
+\bigl|\mu(B_\mc{X}(x,\varepsilon))  - \omega_m \varepsilon^m\bigr|  \leq CmK_0\varepsilon^{m + 2},
+\end{equation*}
+where we recall $K_0$ is an upper bound on the sectional curvature of $\mc{X}$. Plugging this back into~\eqref{pf:error_euclidean_distance_1}, we conclude that
+\begin{align*}
+\int_{\mc{X}} \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') & \leq  \frac{8L_{\eta}\varepsilon^2}{R^2}\Bigl[\omega_m\varepsilon^{m} + CmK_0\varepsilon^{m + 2}\Bigr] + \|\eta\|_{\infty} \Bigl[\omega_m(\wt{\varepsilon}^m - {\varepsilon}^m) + 2CmK_0\varepsilon^{m + 2}\Bigr] \\
+& \leq \frac{8L_{\eta}\varepsilon^2}{R^2}\Bigl[\omega_m\varepsilon^{m} + R^2CmK_0\varepsilon^{m}\Bigr] + \|\eta\|_{\infty} \varepsilon^{m + 2}\Bigl[\frac{27\omega_m}{R^2} + 2CmK_0\Bigr] \\
 & =: C_6 \varepsilon^{m + 2}.
@@
 We begin with a pointwise estimate of $\wt{L}_{P,\varepsilon}f$, facilitated by expressing $w(v) = f(\exp_x(v))$ and $q(v) = p(\exp_x(v))$ in normal coordinates, as in \citep{calder2019}.
-Letting $J_x(v)$ be the Jacobian of the exponential map $\exp_x: B(0,\varepsilon) \subseteq T_x(\mc{X}) \to B_{\mc{X}}(x,\varepsilon)$ evaluated at $v \in T_x(\mc{X})$, we have
+Let $J_x(\cdot)$ be the Jacobian of the exponential map $\exp_x$, we have
 \begin{align*}
@@
 \end{equation*}
-\textcolor{red}{(TODO): This is correct, but feels unacceptably vague.
-On the other hand, I don't want to have to redo the entire proof from the flat Euclidean section, since all the steps will be basically identical.
-Ask Ryan + Siva for guidance.}
 
@@
 \begin{equation*}
-|J_x(v) - 1| \leq CmK\varepsilon^2, \quad  \textrm{for all} ~~ v \in B(0,\varepsilon) \subseteq T_x(\mc{X}).
+|J_x(v) - 1| \leq CmK_0\varepsilon^2, \quad  \textrm{for all} ~~ v \in B(0,\varepsilon) \subseteq T_x(\mc{X}).
 \end{equation*}
@@
 \begin{align*}
-\|A_2\|_{\Leb^2(\mc{X})}^2 & \leq Cm^2K^2 \biggl[\int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr)^2 \eta(\|v\|) q(\varepsilon v) \,dv\biggr] \cdot \biggl[\int_{B(0,1)} \eta(\|v\|) q(\varepsilon v) \,dv\biggr] \\
-& \leq Cm^2K^2 \sigma_{\eta} (1 + L_q\varepsilon)  \int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr)^2 \eta(\|v\|) q(\varepsilon v) \,dv \\
-& \leq Cm^2K^2 \sigma_{\eta}^2 (1 + L_q\varepsilon) p_{\max}  \varepsilon^2 |f|_{H^1(\mc{X})}^2,
+\|A_2\|_{\Leb^2(\mc{X})}^2 & \leq Cm^2K_0^2 \biggl[\int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr)^2 \eta(\|v\|) q(\varepsilon v) \,dv\biggr] \cdot \biggl[\int_{B(0,1)} \eta(\|v\|) q(\varepsilon v) \,dv\biggr] \\
+& \leq Cm^2K_0^2 \sigma_{\eta} (1 + L_q\varepsilon)  \int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr)^2 \eta(\|v\|) q(\varepsilon v) \,dv \\
+& \leq Cm^2K_0^2 \sigma_{\eta}^2 (1 + L_q\varepsilon) p_{\max}  \varepsilon^2 |f|_{H^1(\mc{X})}^2,
 \end{align*}
@@
 \begin{equation*}
-\int_{\mc{X}} \int_{\mc{X}} \bigl(f(x') - f(x)\bigr)^2 \1\{\varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon}\} \,d\mu(x') \,d\mu(x) \leq (1 + CmK\varepsilon^2) \cdot \int_{\mc{X}} \int_{B_{m}(0,\wt{\varepsilon})} |d_x^{1}f(v)|^2 \,dv \,d\mu(x).
+\int_{\mc{X}} \int_{\mc{X}} \bigl(f(x') - f(x)\bigr)^2 \1\{\varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon}\} \,d\mu(x') \,d\mu(x) \leq (1 + CmK_0\varepsilon^2) \cdot \int_{\mc{X}} \int_{B_{m}(0,\wt{\varepsilon})} |d_x^{1}f(v)|^2 \,dv \,d\mu(x).
 \end{equation*}
@@
 \paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_manifold}.}
-Follows identically to the proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~\ref{prop:graph_seminorm_manifold} rather than Proposition~\ref{prop:graph_seminorm_fo}, Proposition~\textcolor{red}{(?)} rather than Proposition~\ref{prop:graph_eigenvalue}, and Assumption~\ref{asmp:parameters_estimation_manifold} rather than Assumption~\ref{asmp:parameters_testing_fo}.
+Follows identically to the proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~\ref{prop:graph_seminorm_manifold} rather than Proposition~\ref{prop:graph_seminorm_fo}, Proposition~\ref{prop:graph_eigenvalue_manifold} rather than Proposition~\ref{prop:graph_eigenvalue}, and Assumption~\ref{asmp:parameters_estimation_manifold} rather than Assumption~\ref{asmp:parameters_testing_fo}.
 
@@
 
-\subsection{Comparison of Laplacian eigenmaps to kernel smoothing}
+\subsection{Comparison of Laplacian Eigenmaps to kernel smoothing}
 \label{subsec:eigenmaps_beats_kernel_smoothing}
@@
 where $\theta \in \Reals$.
-Note that as $n \to \infty$, the density $p^{(n)}$ is converging (in, say, $L^2([0,1])$ norm) to uniform distribution on $[0,1]$, and the regression function $f_0^{(n)}$ to a step function.
+Note that as $n \to \infty$, the density $p^{(n)}$ is converging (in, say, $L^2([0,1])$ norm) to the uniform distribution on $[0,1]$, and the regression function $f_0^{(n)}$ to a step function.
 
@@
 Such a result is available to us because the kernel $\psi$ is Lipschitz on its support, so that the class of functions $\{\psi((x - \cdot)/h): h \in \Rd\}$ has finite VC dimension.
-The precise estimate we use is due to \textcolor{blue}{(Gine + Guillou 2002)}.
+The precise estimate we use is due to~\cite{gine2002}.
 \begin{lemma}[Uniform bound for empirical degree.]
@@
 
-\textcolor{red}{(TODO): Check with Ryan and Siva that (i) they believe these facts, and (ii) these facts do not require additional justification.}
-
-\subsection{Upper bound on $L^2(P_n)$ risk}
-\label{subsec:upper_bound_l2pn_risk}
-\textcolor{red}{(TODO)}
+\section{Computational considerations}
+\label{sec:computational_considerations}
+Recall that when $s = 1$, we have shown that Laplacian eigenmaps is optimal when $\varepsilon \asymp (\log n/n)^{1/d}$ is (up to a constant) as small as possible while still ensuring the graph $G$ is connected.
+On the other hand, when $s > 1$, we can show Laplacian eigenmaps is optimal only when $\varepsilon = \omega(n^{-c})$ for some $c < 1/d$.
+For such a choice of $\varepsilon$, the average degree in $G$ will grow polynomially in $n$ as $n \to \infty$, and computing eigenvectors of the Laplacian of a graph will be more computationally intensive than if the graph were sparse.
+In this dense-graph setting, we now discuss a procedure to more efficiently compute an approximation to the Laplacian eigenmaps estimate: \emph{edge sparsification}.
+
+By now there exist various methods see (e.g., the seminal papers of \citet{spielman2011,spielman2013,spielman2014}, or the overview by \citet{vishnoi2012} and references therein) to efficiently remove many edges from the graph $G$ while only slightly perturbing the spectrum of the Laplacian.
+Specifically such algorithms take as input a parameter $\sigma \geq 1$, and return a sparser graph $\wc{G}$, $E(\wc{G}) \subseteq E(G)$, with a Laplacian $\wc{L}_{n,\varepsilon}$ satisfying
+\begin{equation*}
+\frac{1}{\sigma} \cdot u^{\top} \wc{L}_{n,\varepsilon} u \leq u^{\top} L_{n,\varepsilon} u \leq \sigma \cdot u^{\top} \wc{L}_{n,\varepsilon}u \quad \textrm{for all $u \in \Reals^n$.}
+\end{equation*}
+Let $\wc{f}$ be the Laplacian eigenmaps estimator computed using the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ .
+Because $\wc{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wc{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$, and consequently much faster to compute $\wc{f}$ than $\wh{f}$.
+Statistically speaking, letting $\wc{\lambda}_k$ be the $k$th eigenvalue of $\wc{L}_{n,\varepsilon}$, we have that conditional on $\{X_1,\ldots,X_n\}$,
+\begin{equation*}
+\|\wc{f} - f_0\|_n^2 \leq \frac{\dotp{\wc{L}_{n,\varepsilon}^s f_0}{f_0}_n}{\wc{\lambda}_{K + 1}^s} + \frac{5K}{n} \leq \sigma^{2s} \frac{\dotp{\wc{L}_{n,\varepsilon}^s f_0}{f_0}_n}{\wc{\lambda}_{K + 1}^s} + \frac{5K}{n},
+\end{equation*}
+with probability at least $1 - \exp(-K)$.
+Consequently $\wt{f}$ has $L^2(P_n)$-error of at most $\sigma^{2s}$ times our upper bound on the error of $\wh{f}$, and for any choice of $\sigma$ that is constant in $n$ the estimator $\wc{f}$ will also be rate-optimal.
+
+In fact the aforementioned edge sparsification algorithms are overkill for our needs.
+For one thing, they are designed to work when $\sigma$ is very close to $1$, whereas in order for $\wc{f}$ to be rate-optimal, setting $\sigma$ to be any constant greater than $1$, say $\sigma = 2$, is sufficient.
+Additionally, edge sparsification algorithms are traditionally designed to work in the worst-case, where no assumptions are made on the structure of the graph $G$.
+But the geometric graphs we consider in this paper exhibit a special structure, in which very roughly speaking no single edge is a bottleneck.
+As pointed out by~\citet{sadhanala16b}, in this special case there are far simpler and faster methods for sparsification, which at least empirically seem to do the job.

### 2021-08-31 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Made many small changes to Laplacian Eigenmaps paper. Also converted format back to something Siva would prefer.

@@
 
-\input{sections/introduction}
-\input{sections/preliminaries}
+\section{Introduction}
+\label{sec:introduction}
+Two fundamental statistical goals are to learn and leverage structure of an unknown distribution on the basis of samples.
+Neighborhood graph Laplacians, and their associated spectra, have proven to be powerful and general tools for both purposes.
+They have been successfully used for various statistical tasks such as (spectral) clustering, manifold learning, level-set estimation, semi-supervised learning, etc.
+
+By now there exists a rich literature \citep{koltchinskii2000,belkin07,vonluxburg2008,burago2014,shi2015,singer2017,garciatrillos18,trillos2019, calder2019, cheng2021,dunson2021} explaining this practical success from a theoretical perspective.
+Loosely speaking, these works justify graph spectral methods by showing that the eigenvectors and eigenvalues of graph Laplacians are empirical approximations of population-level objects: the eigenvalues and eigenfunctions of a density-weighted Laplacian operator.
+These population-level objects in turn characterize various structural aspects of the density, such as its high- and low-density regions, the shape and intrinsic dimension of its support, and so forth.
+Together, these insights give a reasonably complete explanation of how graph spectra \emph{learn} the structure of an underlying distribution.
+In contrast, in this paper we develop theory showing that graph spectral methods can effectively \emph{leverage} learned structure to perform downstream (supervised) learning tasks.
+Our main goal is to show that spectral projection procedures---which use eigenvectors of a neighborhood graph Laplacian as features in linear regression---are statistically optimal methods for two problems in nonparametric regression: estimation and goodness-of-fit testing.
+
+The regression methods we consider are based on \emph{Laplacian eigenmaps}, first introduced by~\cite{belkin03a}.\footnote{Technically speaking, as used by~\cite{belkin03a} ``Laplacian eigenmaps'' refers to a spectral embedding procedure.
+What we call Laplacian eigenmaps is a two-step regression method: the first step is this spectral embedding, and the second step computes the estimator $\wh{f}$ by running ordinary least squares in embedding space.
+Since in this work we only ever consider the embedding within the context of this two-step regression method, we take the liberty of referring to the entire procedure as Laplacian eigenmaps.} Given samples $(X_1,Y_1),\ldots,(X_n,Y_n)$, Laplacian eigenmaps first forms a neighborhood graph, with vertices at the design points $\{X_1,\ldots,X_n\}$ and weighted edges between pairs of sufficiently close points $X_i$ and $X_j$.
+The Laplacian eigenmaps estimator $\wh{f}$ is then computed using the spectral decomposition of an (unweighted) Laplacian defined over this neighborhood graph, by projecting the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto a subspace spanned by a certain number of ``low-frequency'' graph Laplacian eigenvectors; more precisely, the eigenvectors corresponding to the smallest eigenvalues of the graph Laplacian.
+The Laplacian eigenmaps test statistic $\wh{T} = \|\wh{f}\|_n^2$ can in turn be used as a statistic to detect whether any signal is present, that is, whether the regression function $f_0(x) = \mathbb{E}[Y|X = x]$ is anywhere non-zero.
+(The precise definition of these methods is given in Section~\ref{subsec:laplacian_eigenmaps}.) We study these methods through the lens of a minimax analysis.
+Our major contribution is to provide upper bounds on the error of the Laplacian eigenmaps estimator and test which imply that they achieve the minimax optimal rates of convergence over natural smoothness classes.
+
+More specifically, we consider the worst-case risk of Laplacian eigenmaps over normed balls in Sobolev spaces.
+Roughly speaking, for a domain $\mc{X} \subseteq \Rd$, the order-$s$ (Hilbert-)Sobolev space $H^s(\mc{X})$ consists of those functions which are $s$-times weakly differentiable, with order-$1$ up through order-$s$ derivatives bounded in $L^2(\mc{X})$ norm.
+Equivalently, Sobolev spaces can be defined in a spectral manner;  letting $\psi_1,\psi_2,\ldots$ be the eigenfunctions of a Laplacian operator defined with respect to $\mc{X}$, Sobolev spaces consist of those functions $f$ whose Fourier coefficients $\theta_k := \int_{\mc{X}} f \psi_k$ are bounded in a particular sequence norm (see Section~\ref{subsec:spectral_projection}).
+The minimax estimation and testing rates for Sobolev spaces are generally speaking well-known---when using mean-squared error as the loss, the estimation rate is $n^{-2s/(2s + d)}$, and the testing rate is $n^{-4s/(4s + d)}$---and are achieved by various methods.
+However, since Sobolev spaces admit a spectral characterization in terms of $\psi_1,\psi_2,\ldots$, spectral projection methods that use these eigenfunctions as features are a particularly natural choice for regression over Sobolev spaces, and are in a sense canonical estimators for these function classes.
+% (AG 8/13/21): Last clause makes a strong statement that I personally believe, but may irritate certain readers. Keep or remove?
+% (AG 8/18/21): I do not love the name ``classical spectral projection method''. It doesn't roll off the tongue, and ``classical'' doesn't really convey what the method is. Is there a better name?
+
+Laplacian eigenmaps serves as a data-dependent approximation to these population-level spectral projection methods.\footnote{We use the term ``population-level'' to emphasize that these methods use eigenfunctions which are the infinite-data limits of graph Laplacian eigenvectors.} This data dependency can be very useful.
+As the name suggests, population-level spectral methods require a priori knowledge of the design distribution, since they are defined with respect to the eigenfunctions of a Laplacian operator which itself depends on this distribution.
+For this reason, these methods are typically studied only under restrictive conditions on the design; for instance, that the design points are uniformly distributed over the unit cube.
+In many situations such restrictive conditions are not satisfied, and it may not be reasonable to assume prior knowledge of the design distribution.
+In contrast to population-level methods, Laplacian eigenmaps implicitly learns structure from the design points $\{X_1,\ldots,X_n\}$, and can thus be used even when the design distribution is unknown.
+Of course, this comes with a trade off.
+Intuitively, using an empirical approximation to the ``right'' population-level basis will incur some additional error.
+The question is whether, notwithstanding this additional error, Laplacian eigenmaps shares the strong statistical properties of its population-level counterpart.
+
+We find that in many cases, the answer is yes: under appropriate regularity conditions, when correctly tuned Laplacian eigenmaps is minimax optimal over Sobolev spaces.
+Specifically, our main contributions are as follows:
+\begin{itemize}
+	\item Over the Sobolev space $H^{s}(\mc{X})$, we establish that the Laplacian eigenmaps estimator $\wh{f}$ has in-sample mean-squared error of at most on the order of $n^{-2s/(2s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d$.
+A test based on the statistic $\|\wh{f}\|_n^2$ has a squared critical radius on the order of $n^{-4s/(4s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d \in \{1,2,3,4\}$.
+	\item We also consider the behavior of Laplacian eigenmaps when the data satisfies a \emph{manifold hypothesis}, meaning the design points $\{X_1,\ldots,X_n\}$ belong to some low-dimensional submanifold of $\Rd$.
+In this case, it is known that the minimax rates over Sobolev spaces depend on the intrinsic dimension $m$ of the manifold: explicitly, the estimation rate is $n^{-2s/(2s + m)}$ and the testing rate is $n^{-4s/(4s + m)}$.
+We give upper bounds showing that Laplacian eigenmaps achieves these rates so long as $s \in \{1,2,3\}$.
+One can interpret these results as precisely quantifying a way in which Laplacian eigenmaps leverages structure, by adapting to the intrinsic dimension of the design.
+	\item The estimator $\wh{f}$ is defined only in-sample.
+We propose a simple out-of-sample extension based on kernel smoothing, and show that this extension adds negligible error, resulting in an estimator which is defined on all of $\mc{X}$ and has out-of-sample mean-squared error also on the order of $n^{-2s/(2s + d)}$.
+\end{itemize}
+In all cases, our bounds also depend optimally on the radius $M$ of the Sobolev ball under consideration.
+
+To derive these results, we develop a framework for analysis reminiscent of that  classically used to analyze population-level spectral projection estimators, but ultimately quite different.
+The chief difference is in the approximation error (bias).
+For our purposes we must establish that with high probability, the eigenvectors of graph Laplacians (which are data-dependent and hence random objects) can be used to effectively approximate continuum Sobolev functions, whereas in the classical analysis an upper bound on the approximation error follows immediately from the aforementioned spectral characterization of Sobolev spaces.
+To tightly upper bound the approximation error of Laplacian eigenmaps, we rely on several recent results regarding concentration of the spectra of graph Laplacians around their population limits, and prove some additional results of this type which may be of independent interest.
+
+For some values of $s$ (number of derivatives) and $d$ (dimension), there do exist gaps between our upper bounds on the error of Laplacian eigenmaps and the minimax rates.
+Although we do not give corresponding lower bounds verifying the tightness of our analysis, we believe these gaps reflect the true behavior of the method rather than some looseness in our analysis, and we comment more on this at relevant parts in the text.
+For completeness, we summarize all of our upper bounds---those which match the minimax rates, and those which do not---in Tables~\ref{tbl:estimation_rates} and~\ref{tbl:testing_rates}.
+\begin{table}
+	\begin{center}
+		\begin{tabular}{p{.2\textwidth} | p{.14\textwidth} p{.12\textwidth} }
+			Smoothness order & Flat Euclidean (Model~\ref{def:model_flat_euclidean}) & Manifold (Model~\ref{def:model_manifold}) \\
+			\hline
+			$s \leq 3$ & $\bm{n^{-2s/(2s + d)}}$ & $\bm{n^{-2s/(2s + m)}}$ \\
+			$s > 3$  & $\bm{n^{-2s/(2s + d)}}$ & $n^{-6/(6 + m)}$
+		\end{tabular}
+	\end{center}
+	\caption{Summary of Laplacian eigenmaps estimation rates over Sobolev balls.
+Bold font marks minimax optimal rates.
+In each case, rates hold for all $d \in \mathbb{N}$ (under Model~\ref{def:model_flat_euclidean}), and for all $m \in \mathbb{N}, 1 < m < d$ (under Model~\ref{def:model_manifold}).
+Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps estimator is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal}
+	\label{tbl:estimation_rates}
+\end{table}
+
+\begin{table}
+	\begin{center}
+		\begin{tabular}{p{.175\textwidth} p{.175\textwidth} | p{.14\textwidth} p{.12\textwidth} }
+			Smoothness order & Dimension & Flat Euclidean (Model~\ref{def:model_flat_euclidean}) & Manifold (Model~\ref{def:model_manifold}) \\
+			\hline
+			\multirow{2}{*}{$s = 1$} & $\dim(\mc{X}) < 4$ & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
+			& $\dim(\mc{X}) \geq 4$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
+			\hline
+			\multirow{3}{*}{$s = 2$ or $3$} & $\dim(\mc{X}) \leq 4$  & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
+			& $4 <\dim(\mc{X}) < 4s$  & $n^{-2s/(2(s - 1) + d)}$ & $n^{-2s/(2(s - 1) + m)}$\\
+			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
+			\hline
+			\multirow{3}{*}{$s > 3$} & $\dim(\mc{X}) \leq 4$ & $\bm{n^{-4s/(4s + d)}}$ & $n^{-12/(12 + d)}$ \\
+			& $4 < \dim(\mc{X}) < 4s$ & $n^{-2s/(2(s - 1) + d)}$ & $n^{-6/(4 + m)}$ \\
+			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
+		\end{tabular}
+	\end{center}
+	\caption{Summary of Laplacian eigenmaps testing rates over Sobolev balls.
+Bold font marks minimax optimal rates.
+Rates when $d > 4s$ assume that $f_0 \in L^4(\mc{X})$, and depend on $\|f_0\|_{L^4(\mc{X})}$.
+Although we suppress it for simplicity, in all cases when othe Laplacian eigenmaps test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal.}
+	\label{tbl:testing_rates}
+\end{table}
+
+\paragraph{Related work.}
+
+Most work on supervised learning using graphs adopts a \emph{fixed design} perspective, treating the design points $X_1 = x_1,\ldots,X_n = x_n$ as vertices of a fixed graph, and carrying out inference with respect to the conditional mean vector $(f_0(x_1),\ldots,f_0(x_n))$.
+In this setting, matching upper and lower bounds have been established that certify the optimality of graph-based methods for estimation \citep{wang2016,hutter2016,sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}) and testing \citep{sharpnack2010identifying,sharpnack2013b,sharpnack2013,sharpnack2015} over different ``function'' classes (in quotes because these classes really model the $n$-dimensional vector of evaluations).
+This setting is quite general, because the graph need not be a geometric graph defined on a vertex set which belongs to Euclidean space.
+On the other hand, depending on the data collection process, it may be unnatural to model the design points as being a priori fixed, and the estimand as being a vector which exhibits a discrete notion of ``smoothness'' over this fixed design.
+Instead, we adopt the \emph{random design} perspective, and seek to estimate a function that we assume exhibits a more classical notion of smoothness.
+
+In the random design setting the analysis of graph-based approaches to supervised learning is more challenging, and only a few works have studied the matter.
+\cite{zhou2011} provide an asymptotic analysis of the Laplacian eigenmaps estimator in the semi-supervised setting, as the number of unlabeled samples tends to infinity; in this case the estimator is equivalent to population-level spectral projection, and there is no extra error incurred by using data-dependent eigenvectors.
+\cite{lee2016} consider the \emph{diffusion maps} estimator---which uses the eigenvectors of a differently normalized graph Laplacian---in a similar supervised setting to our own, but prove suboptimal rates of convergence.
+
+Finally, \citet{trillos2020,green2021} study \emph{graph Laplacian regularization}, a graph-based penalized least squares method which uses a quadratic form involving the graph Laplacian as a penalty.
+Laplacian regularization can, like Laplacian eigenmaps, be cast as a spectral method.
+The difference is that Laplacian regularization induces some smooth shrinkage based on the eigenvalue, whereas Laplacian eigenmaps corresponds to a hard spectral cutoff.
+Interestingly, this difference leads to a stark contrast in the ultimate statistical properties of the two approaches.
+For instance, in~\cite{green2021} Laplacian regularization is shown to be optimal only for the first-order Sobolev space $H^1(\mc{X})$, and even then only when $d \in \{1,2,3,4\}$.
+In contrast, in this work we show that the Laplacian eigenmaps estimator is optimal for all $s$ and $d$.
+
+\paragraph{Roadmap.}
+We now outline the structure of the rest of this paper.
+In Section~\ref{sec:setup_main_results}, we give our formal modeling assumptions, and precisely define the Laplacian eigenmaps methods we study.
+Propositions~\ref{prop:spectral_series_estimation} and ~\ref{prop:spectral_series_testing}, in Section~\ref{subsec:spectral_projection}, show that under rather general (nonparametric) conditions on the design distribution, population-level spectral projection methods achieve minimax rates of convergence over Sobolev classes.
+Then in Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps},~\ref{sec:manifold_adaptivity}, and \ref{sec:out_of_sample}, we give our main upper bounds on the error of Laplacian eigenmaps (summarized above); these upper bounds hold under similarly general conditions, and imply that Laplacian eigenmaps estimators and tests are also minimax rate-optimal.
+In Section~\ref{sec:experiments} we examine the empirical behavior of Laplacian eigenmaps, and show that even at moderate sample sizes Laplacian eigenmaps is competitive with population-level spectral projection.
+We conclude with some discussion in Section~\ref{sec:discussion}.
+
+\paragraph{Notation.}
+We frequently refer to various classical function classes.
+For an open set $\mc{X}$ equipped with a volume form $d\mu$, we let $L^2(\mc{X})$ denote the set of functions $f$ for which $\|f\|_{L^2(\mc{X})}^2 := \int f^2 \,d\mu  < \infty$, and equip $L^2(\mc{X})$ with the norm $\|\cdot\|_{L^2(\mc{X})}$.
+We define $\dotp{f}{g}_P := \int fg\,dP$, and let $L^2(P)$ contain those functions $f$ for which $\|f\|_P^2 := \dotp{f}{f}_P$ is finite.
+Equivalently, we let $L^2(P_n)$ consist of those ``functions'' $f: \set{X_1,\ldots,X_n} \to \Reals$ for which the empirical norm $\|f\|_{n}^2 := \frac{1}{n}\sum_{i = 1}^{n} \bigl(f(X_i)\bigr)^2 < \infty$.
+When there is no chance of confusion, we will sometimes associate functions in $L^2(P_n)$ with vectors in $\Reals^n$, and vice versa.
+We use $C^k(\mc{X})$ to refer to functions which are $k$ times continuously differentiable in $\mc{X}$, either for some integer $k \geq 1$ or for $k = \infty$.
+We let $C_c^{\infty}(\mc{X})$ represent those functions in $C^{\infty}(\mc{X})$ with support $V$ compactly contained in $\mc{X}$, meaning $\wb{V}$ is compact and $\wb{V} \subseteq \mc{X}$.
+We write $\partial f/\partial r_i$ for the partial derivative of $f$ in the $i$th standard coordinate of $\Rd$, and use the multi-index notation $D^{\alpha}f := \partial^{|\alpha|}f/\partial^{\alpha_1}x_1\ldots\partial^{\alpha_d}x_d$ for multi-indices $\alpha \in \Reals^d$.
+Recall that for a given multi-index $\alpha \in \mathbb{N}^d$, a function $f$ is \emph{$\alpha$-weakly differentiable} if there exists some $h \in L^1(\mc{X})$ such that
+\begin{equation*}
+\int_{\mc{X}} h g = (-1)^{|\alpha|} \int_{\mc{X}} f D^{\alpha}g, \quad \textrm{for every $g \in C_c^{\infty}(\mc{X})$.}
+\end{equation*}
+If such a function $h$ exists, it is the $\alpha$th weak partial derivative of $f$, and denoted by $D^{\alpha}f := h$.
+For functions $f$ which are $|\alpha|$-times classically differentiable, this coincides with the classical definition of derivative, and so we use the same notation for both.
+
+We write $\|\cdot\| = \|\cdot\|_2$ for Euclidean norm, $|\cdot| = \|\cdot\|_1$ for $\ell_1$ norm, and $d_{\mc{X}}(x',x)$ for the geodesic distance between points $x$ and $x'$ on a manifold $\mc{X}$.
+Then for a given $\delta > 0$, $B(x,\delta)$ is the radius-$\delta$ ball with respect to Euclidean distance, whereas $B_{\mc{X}}(x,\delta)$ is the radius-$\delta$ ball with respect to geodesic distance.
+Letting $T_x(\mc{X})$ be the tangent space at a point $x \in \mc{X}$, we write $B_m(v,\delta) \subset T_x(\mc{X})$ for the radius-$\delta$ ball centered at $v \in T_x(\mc{X})$.
+
+For sequences $(a_n)$ and $(b_n)$, we use the asymptotic notation $a_n \lesssim b_n$ to mean that there exists a number $C$ such that $a_n \leq C b_n$ for all $n$.
+We write $a_n \asymp b_n$ when $a_n \lesssim b_n$ and $b_n \lesssim a_n$.
+On the other hand we write $a_n = o(b_n)$ when $\lim a_n/b_n = 0$, and likewise $a_n = \omega(b_n)$ when $\lim a_n/b_n = \infty$.
+Finally $a \vee b := \max\{a,b\}$ and $a \wedge b := \min\{a,b\}$.
+
+\section{Preliminaries}
+\label{sec:setup_main_results}
+
+We begin in Sections~\ref{subsec:regression_laplacian_eigenmaps}-\ref{subsec:laplacian_eigenmaps} by precisely defining the models (random design points, Sobolev regression functions) and methods (Laplacian eigenmaps) under consideration.
+Then in Section~\ref{subsec:spectral_projection}, we connect Sobolev spaces to Laplacian eigenmaps using the spectrum of a density-dependent Laplacian operator, and show in Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} that projection methods which use the eigenfunctions of this operator are statistically optimal.
+
+\subsection{Nonparametric regression over Sobolev spaces}
+\label{subsec:regression_laplacian_eigenmaps}
+
+We will always operate in the usual setting of nonparametric regression with random design.
+We observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$; the design points $X_1,\ldots,X_n$ are sampled from a distribution $P$ with support $\mc{X} \subseteq \Rd$, and the responses follow the signal plus noise model
+\begin{equation}
+\label{eqn:model}
+Y_i = f_0(X_i) + w_i,
+\end{equation}
+with regression function $f_0: \mc{X} \to \Reals$, and $w_i \sim N(0,1)$ independent Gaussian noise.
+% (AG 8/4/21): Removed the comment regarding noise with variance = \sigma^2.
+
+We now formulate two models, which differ in the assumed nature of the support $\mc{X}$ of the design distribution $P$: the \emph{flat Euclidean} and \emph{manifold} models.
+
+\paragraph{Flat Euclidean model.}
+In Definitions~\ref{def:model_flat_euclidean}-\ref{def:zero_trace_sobolev_space}, we collect the assumptions we make when working under the flat Euclidean model.
+We begin by giving some regularity conditions on the design.
+
+\begin{definition}[Flat Euclidean model]
+	\label{def:model_flat_euclidean}
+	The support $\mc{X}$ of the design distribution $P$ is an open, connected, and bounded subset of $\Rd$, with Lipschitz boundary.
+The distribution $P$ admits a Lipschitz density $p$ with respect to the $d$-dimensional Lebesgue measure $\nu$, which is bounded away from $0$ and $\infty$,
+	\begin{equation*}
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty, \quad \textrm{for all $x \in \mc{X}$.}
+	\end{equation*}
+\end{definition}
+% (AG 8/13/21): I got rid of the sentence ``The data are sampled according to (1).'' It seemed repetitive. 
+At various points we will also assume that the density $p \in C^k(\mc{X})$.
+On the other hand, we model the regression function as belonging to an order-$s$ Sobolev space, and being bounded in Sobolev norm.
+\begin{definition}[Sobolev space on a flat Euclidean domain]
+	\label{def:sobolev_space}
+	For an integer $s \geq 1$, a function $f \in L^2(\mc{X})$ belongs to the Sobolev space $H^s(\mc{X})$ if for all $|\alpha| \leq s$, the weak derivatives $D^{\alpha}f$ exist and satisfy $D^{\alpha}f \in L^2(\mc{X})$.
+The $j$th order semi-norm for $f \in H^s(\mc{X})$ is $|f|_{H^j(\mc{X})} := \sum_{|\alpha| = j}\|D^{\alpha}f\|_{\Leb^2(\mc{X})}$, and the corresponding norm
+	\begin{equation*}
+	\|f\|_{H^s(\mc{X})}^2 := \|f\|_{\Leb^2(\mc{X})}^2 + \sum_{j = 1}^{s} |f|_{H^j(\mc{X})}^2,
+	\end{equation*}
+	induces the Sobolev ball
+	\begin{equation*}
+	H^s(\mc{X};M) := \bigl\{f \in H^s(\mc{X}): \|f\|_{H^s(\mc{X})} \leq M\bigr\}.
+	\end{equation*} 
+\end{definition}
+When $s > 1$ we will also assume that $f_0$ satisfies a zero-trace boundary condition.
+Recall that $H^s(\mc{X})$ can alternatively be defined as the completion of $C^{\infty}(\mc{X})$ in the Sobolev norm $\|\cdot\|_{H^s(\mc{X})}$.
+The zero-trace Sobolev spaces are defined in a similar fashion, as the completion of $C_c^{\infty}(\mc{X})$ in the same norm.
+
+\begin{definition}[Zero-trace Sobolev space]
+	\label{def:zero_trace_sobolev_space}
+	A function $f \in H^s(\mc{X})$ belongs to the zero-trace Sobolev space $H_0^s(\mc{X})$ if there exists a sequence $f_1,f_2,\ldots$ of functions in $C_c^{\infty}(\mc{X})$ such that
+	\begin{equation*}
+	\lim_{k \to \infty}\|f_k - f\|_{H^s(\mc{X})} = 0.
+	\end{equation*}
+	The normed ball $H_0^{s}(\mc{X};M) := H_0^{s}(\mc{X}) \cap H^{s}(\mc{X};M)$.
+\end{definition}
+Boundary conditions play an important role in the analysis of spectral methods, as we explain further in Section~\ref{subsec:spectral_projection}.
+For now, we limit ourselves to pointing out that for functions $f \in C^\infty(\mc{X})$, the zero-trace condition can be stated more concretely, as implying that $\partial^{k}f/\partial{\bf n}^k(x) = 0$ for each $k = 0,\ldots,s - 1$, and for all $x \in \partial\mc{X}$.
+(Here $\partial/(\partial {\bf n})$ is the partial derivative operator in the direction of the normal vector $\mathbf{n}$.)
+\paragraph{Manifold model.}
+As in the flat Euclidean case, we start with some regularity conditions on the design. One such condition will be on the reach $R$ of the manifold $\mc{X}$, which we recall is defined as follows:
+\begin{equation*}
+R := \Bigl\{\sup_{r > 0}: \forall z \in \Rd, \inf_{x \in \mc{X}} \|z - x\| \leq r, ~\exists! y \in \mc{X}~\mathrm{s.t.}~\|z - y\| = \inf_{x \in \mc{X}} \|z - x\|\Bigr\}.
+\end{equation*}
+In words, the reach is the largest radius of a ball which can be rolled around the manifold $\mc{X}$.
+\begin{definition}[Manifold model]
+	\label{def:model_manifold}
+	The support $\mc{X}$ of the design distribution $P$ is a closed, connected, and smooth Riemannian manifold (without boundary) embedded in $\Rd$, of intrinsic dimension $1 \leq m < d$, and with a positive reach $R > 0$.
+The design distribution $P$ admits a Lipschitz density $p$ with respect to the volume form $d\mu$ induced by the Riemannian structure of $\mc{X}$, which is bounded away from $0$ and $\infty$,
+	\begin{equation*}
+	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty, \quad \textrm{for all $x \in \mc{X}$.}
+	\end{equation*}
+\end{definition}
+
+There are several equivalent ways to define Sobolev spaces on smooth Riemannian manifolds.
+We will stick with a definition that parallels our setup in the flat Euclidean setting as much as possible.
+To do so, we first recall the notion of partial derivatives on a manifold, which are defined with respect to a local coordinate system.
+Letting $r_1,\ldots,r_m$ be the standard basis of $\Reals^m$, for a given chart $(\phi,U)$ (meaning an open set $U \subseteq \mc{X}$, and a smooth mapping $\phi: U \to \Reals^m$) we write $\phi =: (x_1,\ldots,x_m)$ in local coordinates, meaning $x_i = r_i \circ \phi$.
+Then we define the partial derivative $\partial f/\partial x_i$ of a function $f: \mc{X} \to \Reals$ at $x \in U$ to be
+\begin{equation*}
+\frac{\partial f}{\partial x_i}(x) := \frac{\partial(f \circ \phi^{-1})}{\partial r_i}\bigl(\phi(x)\bigr).
+\end{equation*}
+The right hand side should be interpreted in the weak sense of derivative.
+As before, we use the multi-index notation $D^{\alpha}f := \partial^{|\alpha|}f/\partial^{\alpha_1}x_1\ldots\partial^{\alpha_m}x_m$.
+
+\begin{definition}[Sobolev space on a manifold]
+	\label{def:sobolev_space_manifold}
+	A function $f \in \Leb^2(\mc{X})$ belongs to the Sobolev space $H^{s}(\mc{X})$ if for all $\abs{\alpha} \leq s$, the weak derivatives $D^{\alpha}f$ exist and satisfy  $D^{\alpha}f \in \Leb^2(\mc{X})$.
+The $j$th order semi-norm $|f|_{H^j(\mc{X})}$, the norm $\|f\|_{H^s(\mc{X})}$, and the ball $H^s(\mc{X};M)$ are all defined as in Definition~\ref{def:sobolev_space}.
+\end{definition}
+The partial derivatives $D^{\alpha}f$ clearly depend on the choice of local coordinates, and so will the resulting Sobolev norm~$\|f\|_{H^s(\mc{X})}$.
+However, for our purposes the important thing is that regardless of the choice of local coordinates the resulting norms will be equivalent\footnote{Recall that norms $\|\cdot\|_1$ and $\|\cdot\|_2$ on a space $\mc{F}$ are said to be equivalent if there exist constants $c$ and $C$ such that
+	\begin{equation*}
+	c \|f\|_1 \leq \|f\|_2 \leq C \|f\|_1 \quad \textrm{for all $f \in \mc{F}$.}
+	\end{equation*}} 
+and so the ultimate Sobolev space $H^s(\mc{X})$ is independent of local coordinates.
+For more information regarding manifolds and Sobolev spaces defined thereupon, see~\cite{lee2013} and~\cite{hebey1996}.
+
+\subsection{Laplacian Eigenmaps}
+\label{subsec:laplacian_eigenmaps}
+We now formally define the estimator and test statistic we study.
+Both are derived from eigenvectors of a graph Laplacian.
+For a positive, symmetric kernel $\eta: [0,\infty) \to [0,\infty)$, and a radius parameter $\varepsilon > 0$, let $G = ([n],W)$ be the neighborhood graph formed over the design points $\{X_1,\ldots,X_n\}$, with a weighted edge $W_{ij} = \eta(\|X_i - X_j\|/\varepsilon)$ between vertices $i$ and $j$.
+Then the
+\emph{neighborhood graph Laplacian} $L_{n,\varepsilon}: \Reals^n \to \Reals$ is defined by its action on vectors $u \in \Reals^n$ as
+\begin{equation}
+\label{eqn:neighborhood_graph_laplacian}
+\bigl(L_{n,\varepsilon}u\bigr)_i := \frac{1}{n\varepsilon^{2 + \mathrm{dim}(\mc{X})}} \sum_{j = 1}^{n} \bigl(u_i - u_j\bigr) \eta\biggl(\frac{\|X_i - X_j\|}{\varepsilon}\biggr).
+\end{equation}
+(Here $\mathrm{dim}(\mc{X})$ stands for the dimension of $\mc{X}$.
+It is equal to $d$ under the assumptions of Model~\ref{def:model_flat_euclidean}, and equal to $m$ under the assumptions of Model~\ref{def:model_manifold}.
+The pre-factor $(n\varepsilon^{2 + \mathrm{dim}(\mc{X})})^{-1}$ ensures non-degenerate stable limits as $n \to \infty, \varepsilon \to 0$).
+Note that $(n\varepsilon^{\dim(\mc{X}) + 2}) \cdot L_{n,\varepsilon} = D - W$, where $D \in \Reals^{n \times n}$ is the diagonal degree matrix, $D_{ii} = \sum_{i = 1}^{n} W_{ij}$.
+
+The graph Laplacian is a positive semi-definite matrix, and admits the eigendecomposition $L_{n,\varepsilon} = \sum_{k = 1}^{n} \lambda_k v_k v_k^{\top}$, where for each $k \in \{1,\ldots,n\}$ the eigenvalue-eigenvector pair $(\lambda_k,v_k)$ satisfies
+\begin{equation*}
+L_{n,\varepsilon}v_k = \lambda_k v_k, \quad \|v_k\|_2^2 = 1.
+\end{equation*}
+We will assume without loss of generality that each eigenvalue $\lambda$ of $L_{n,\varepsilon}$ has algebraic multiplicity $1$, and so we can index the eigenpairs $(\lambda_1,v_1),\ldots,(\lambda_n,v_n)$ in ascending order of eigenvalue, $0 = \lambda_1 < \ldots < \lambda_n$.
+
+The Laplacian eigenmaps estimator $\wh{f}$ simply projects the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto the first $K$ eigenvectors of $L_{n,\varepsilon}$: letting $V_K \in \Reals^{n \times K}$ be the matrix with columns $V_{K,k} = v_k$, we have that
+\begin{equation}
+\label{eqn:laplacian_eigenmaps_estimator}
+\wh{f} := \sum_{k = 1}^{K} \dotp{{\bf Y}}{v_k}_{2} v_k = V_K V_K^{\top} {\bf Y}.
+\end{equation} 
+If $\wh{f}$ is a reasonable estimate of $f_0$, then the Laplacian eigenmaps test statistic,
+\begin{equation}
+\label{eqn:laplacian_eigenmaps_test}
+\wh{T} := \|\wh{f}\|_n^2 = \frac{1}{n} {\bf Y}^{\top} V_K V_K^{\top} {\bf Y},
+\end{equation}
+is in turn a reasonable estimate of $\|f_0\|_{P}^2$, and can be used in the \emph{signal detection} problem to distinguish whether or not $f_0 = 0$.
+
+\subsection{Spectral Projection using Continuum Eigenfunctions}
+\label{subsec:spectral_projection}
+Laplacian eigenmaps serves as a data-dependent alternative to more classical population-level spectral projection methods, which operate by projecting the response vector ${\bf Y}$ onto a subspace spanned by eigenfunctions of some differential operator.
+For instance, projecting ${\bf Y}$ onto the eigenfunctions of the Laplacian operator $\Delta = \sum_{i = 1}^{d} \partial^2f/\partial x_i^2$ corresponds to estimating $f_0$ using noisy empirical Fourier coefficients.
+
+We will focus on a density-dependent alternative to the Laplacian,  $\Delta_P:C^2(\mc{X}) \to L^2(\mc{X})$,
+\begin{equation}
+\label{eqn:laplace_beltrami}
+\Delta_Pf := -\frac{1}{p} \mathrm{div}(p^2\nabla f),
+\end{equation}
+which has enumerable eigenvalue/eigenfunction pairs $(\lambda_k(\Delta_P),\psi_k),(\lambda_2(\Delta_P),\psi_2),\ldots$ defined according to
+\begin{equation}
+\label{eqn:laplace_beltrami_eigenproblem}
+\Delta_P\psi = \lambda(\Delta_P) \psi, \quad \frac{\partial}{\partial{\bf n}}\psi = 0~~\textrm{on $\partial \mc{X}$,}
+\end{equation}
+and sorted as usual in ascending order of eigenvalue.\footnote{For formal justification of the fact that~\eqref{eqn:laplace_beltrami_eigenproblem} has a discrete spectrum under either Model~\ref{def:model_flat_euclidean} or~\ref{def:model_manifold}, see~\cite{garciatrillos18,trillos2019}.} Formally, the classical spectral series estimator and test statistic are\footnote{Technically speaking, $\wt{f}$ is not the projection of ${\bf Y}$ onto the subspace of $L^2(P_n)$ spanned by $\psi_1,\ldots,\psi_K$, since $\dotp{\psi_k}{\psi_{\ell}}_n \neq 0$.
+However, these eigenfunctions are very close to orthogonal in $L^2(P_n)$, orthogonalization will have a negligible impact on the overall statistical properties, and so for simplicity we consider the methods as defined in~\eqref{eqn:classical_spectral_projection}.}
+\begin{equation}
+\label{eqn:classical_spectral_projection}
+\wt{f}(x) := \sum_{k = 1}^{K} \dotp{{\bf Y}}{\psi_k}_{n} \psi_k(x),\quad\textrm{and}\quad \wt{T} := \|\wt{f}\|_P^2.
+\end{equation}
+
+We focus on the particular differential operator $\Delta_P$ because the spectrum of $\Delta_P$ is the population-level limit of the graph Laplacian spectrum: for any fixed $k \in \mathbb{N}$, as the number of samples $n \to \infty$, the graph radius $\varepsilon \to 0$, and $n\varepsilon^{2 + \dim(\mc{X})} \to \infty$,
+\begin{equation}
+\lambda_k \to \lambda_k(\Delta_P), \quad\textrm{and}\quad \Bigl\|\frac{1}{n}v_k - \psi_k\Bigr\|_n^2 \to 0.
+\end{equation}
+This holds in either the flat Euclidean (Model~\ref{def:model_flat_euclidean}) or manifold (Model~\ref{def:model_manifold}) setups~\citep{garciatrillos18,trillos2019}.
+It follows that for any fixed number of eigenvectors $K \in \mathbb{N}$, both $\wh{f} \to \wt{f}$ and $\wh{T} \to \wt{T}$, and in this way Laplacian eigenmaps can be formally tied to more classical spectral projection methods for regression.
+
+On the other hand, the eigenvalues and eigenfunctions of $\Delta_P$ can also be used to give a spectral definition of Sobolev spaces on general domains $\mc{X}$. Consider the ellipsoid
+\begin{equation}
+\label{eqn:sobolev_ellipsoid}
+\mc{H}^{s}(\mc{X}) := \Bigl\{\sum_{k = 1}^{\infty} a_k \psi_k \in L^2(\mc{X}):  \sum_{k = 1}^{\infty} a_k^2 \lambda_{k}^s(\Delta_P) \leq M^2 \Bigr\},
+\end{equation}
+equipped with the norm $\|\sum_{k = 1}^{\infty} a_k \psi_k\|_{\mc{H}^s(\mc{X})}^2 = \sum_{k = 1}^{\infty} a_k^2 \lambda_{k}^s(\Delta_P)$.
+Under appropriate regularity conditions $\mc{H}^s(\mc{X})$ consists of functions $f \in H^s(\mc{X})$ which also satisfy some additional boundary conditions.
+For instance, assuming Model~\ref{def:model_flat_euclidean}, $p \in C^{\infty}(\mc{X})$ and $\partial \mc{X} \in C^{1,1}$, \citet{dunlop2020} show that for any $s \geq 1$, the ellipsoid $\mc{H}^{2s}(\mc{X})$ satisfies
+\begin{equation}
+\label{eqn:sobolev_ellipsoid_to_sobolev_ball}
+\mc{H}^{2s}(\mc{X}) = 
+\biggl\{f \in H^{2s}(\mc{X}): \frac{\partial \Delta_P^rf}{\partial {\bf n}} = 0~\textrm{on}~\partial\mc{X},~~\textrm{for all $0 \leq r \leq s - 1$} \biggr\},
+\end{equation}
+and likewise $\mc{H}^{2s + 1}(\mc{X}) = \mc{H}^{2s}(\mc{X}) \cap H^{2s + 1}(\mc{X})$ for any $s \geq 0$; additionally, the norms $\|\cdot\|_{\mc{H}^s(\mc{X})}$ and $\|\cdot\|_{H^s(\mc{X})}$ are equivalent.
+
+This spectral formulation suggests that population-level spectral projection methods---and in turn Laplacian eigenmaps---are natural choices for regression over Sobolev spaces.
+Furthermore, since these population-level methods serve, in a sense, as the infinite-data limit of Laplacian eigenmaps, the statistical properties of the former---which are generally speaking better understood and easier to derive---should intuitively shed some light on the properties of the latter.
+To this end, we now state a pair of results, Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, which establish that population-level spectral projection methods are statistically optimal for nonparametric estimation and testing over Sobolev ellipsoids.
+These results are generalizations of previously known upper bounds on the estimation and testing error of classical spectral methods~\citep{tsybakov08,ingster2009}, but they hold under less stringent (nonparametric) conditions on the design distribution $P$.
+This latter point is important, since as we have argued one attractive quality of Laplacian eigenmaps is its adaptivity to general design distributions.
+
+\paragraph{Estimation.}
+We begin with an upper bound on the $L^2(P)$ risk of $\wt{f}$.
+\begin{proposition}
+	\label{prop:spectral_series_estimation}
+	Suppose data is observed according to Model~\ref{def:model_flat_euclidean}, and that the density $p$ is known.
+Suppose additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$, $f_0 \in \mc{H}^{s}(\mc{X};M)$ and $\|f_0\|_P^2 \leq 1$.
+Then there exists a constant $C$ which does not depend on $f_0$, $M$ or $n$ such that the following statement holds: if the spectral projection estimator $\wt{f}$ is computed with parameter $K = \max\{\floor{M^2n}^{d/(2s + d)},1\}$, then
+	\begin{equation}
+	\label{eqn:spectral_series_estimation}
+	\Ebb\bigl[\|\wt{f} - f_0\|_P^2\bigr] \leq C \min\Bigl\{M^2\bigl(M^2n\bigr)^{-2s/(2s + d)}, \frac{1}{n}\Bigr\}.
+	\end{equation}
+\end{proposition}
+When the Sobolev ball radius $n^{-1/2} \lesssim M$, the upper bound in~\eqref{eqn:spectral_series_estimation} is on the order of $M^2(M^2n)^{-2s/(2s + d)}$, which is the well-known minimax rate for estimation over Sobolev classes (see e.g.~\cite{gyorfi2006,wasserman2006,tsybakov08} and references therein, and specifically Theorem~3.2 of~\cite{gyorfi2006} for a matching lower bound in the context of nonparametric regression with random design).
+
+% (AG 8/19/21): I would like to strengthen this statement to not require $M \asymp 1$, but I do not know where I can find the rate M^2(M^2n)^{-2s/(2s + d)} for all s and d combinations.
+In fact, I don't even know of a paper that states the M^2(M^2n)^{-2s/(2s + d)} rate for L^2(P) risk in random design.
+
+We now give the proof of Proposition~\ref{prop:spectral_series_estimation}.
+The structure of the analysis, which is fairly classical and straightforward, can be usefully compared to our analysis of Laplacian eigenmaps (see Section~\ref{subsec:analysis}).
+
+\paragraph{Proof of Proposition~\ref{prop:spectral_series_estimation}.}
+We decompose risk into squared bias and variance,
+\begin{equation}
+\label{pf:spectral_series_estimation_0}
+\Ebb \|\wt{f} - f_0\|_P^2 = \Ebb\| \Ebb[\wt{f}]  - f_0\|_P^2 + \Ebb\| \wt{f} - \Ebb[\wt{f}]\|_P^2.
+\end{equation}
+Since the eigenfunctions $\{\psi_k\}$ form an orthonormal basis of $L^2(P)$, and $f_0 \in \mc{H}^{s}(\mc{X}) \subseteq L^2(P)$, we can write the squared bias in terms of squared Fourier coefficients of $f_0$, leading to the following upper bound,
+\begin{equation*}
+\|f_0 - \Ebb[\wt{f}]\|_P^2 = \sum_{k = K + 1}^{\infty}  \dotp{f_0}{\psi_k}_P^2 \leq  \frac{1}{\{\lambda_{K + 1}(\Delta_P)\}^s} \sum_{k = K + 1}^{\infty} \{\lambda_{k + 1}(\Delta_P)\}^s \dotp{f_0}{\psi_k}_P^2 \leq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}}{\{\lambda_{K + 1}(\Delta_P)\}^s}.
+\end{equation*}
+On the other hand, the variance term can be written as the sum of the variance of each empirical Fourier coefficient, and subsequently by the law of total variance we derive that
+\begin{align}
+\label{pf:spectral_series_estimation_2}
+\Ebb\| \wt{f} - \Ebb[\wt{f}]\|_P^2 = \sum_{k = 1}^{K} \Var\Bigl[\dotp{{\bf Y}}{\psi_k}_n\Bigr] & = \sum_{k = 1}^{K} \Var\Bigl[\Ebb[\dotp{Y}{\psi_k}_n|{\bf X}\Bigr] + \Ebb\Bigl[\Var[\dotp{Y}{\psi_k}_n|{\bf X}\Bigr] \nonumber \\
+& = \sum_{k = 1}^{K} \Var\Bigl[\dotp{f_0}{\psi_k}_n\Bigr] + \frac{1}{n}\Ebb\Bigl[\|\psi_k\|_n^2\Bigr] \nonumber \\
+& \leq \frac{K}{n} + \frac{1}{n}\sum_{k = 1}^{K}\Ebb\Bigl[\Bigl(f_0(X)\psi_k(X)\Bigr)^2\Bigr].
+\end{align}
+Consequently,
+\begin{equation}
+\label{pf:spectral_series_estimation_1}
+\Ebb \|\wt{f} - f_0\|_P^2 \leq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\bigl[\lambda_{K + 1}(\Delta_P)\bigr]^s} + \frac{K}{n} + \frac{1}{n}\Ebb\Bigl[(f_0(X))^2 \cdot \sum_{k = 1}^{K} (\psi_k(X))^2\Bigr].
+\end{equation}
+The claim of the proposition then follows from variants of two classical results in spectral geometry.
+The first is a Weyl's Law asymptotic scaling of the eigenvalues of $\Delta_P$ due to~\cite{dunlop2020}; formally, there exist constants $c$ and $C$ (which will depend on $P$ and $d$) such that
+\begin{equation}
+\label{eqn:weyl}
+ck^{2/d} \leq \lambda_k(\Delta_P) \leq Ck^{2/d}\quad\textrm{for all $k \in \mathbb{N}$, $k \geq 2$}.
+\end{equation}
+The second is a local analog to Weyl's Law, which says that there exists a constant $C$ (again depending on $P$ and $d$) such that
+\begin{equation}
+\label{eqn:local_weyl}
+\sup_{x \in \mc{X}}\biggl\{\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2\biggr\} \leq CK \quad\textrm{for all $K \in \mathbb{N}$}.
+\end{equation}
+Equation~\eqref{eqn:local_weyl} is a direct implication of~\eqref{eqn:weyl} along with Theorem 17.5.3 of~\cite{hormander1973}.
+Plugging the upper bounds~\eqref{eqn:weyl} and~\eqref{eqn:local_weyl} back into~\eqref{pf:spectral_series_estimation_1}, we conclude that
+\begin{equation}
+\label{pf:spectral_series_estimation_3}
+\Ebb \|\wt{f} - f_0\|_P^2 \leq C\biggl(\frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{(K + 1)^{2s/d}} + \frac{K}{n}\biggr).
+\end{equation}
+If $n^{-1/2} \geq M$, then taking $K = 1$ implies $\Ebb \|\wt{f} - f_0\|_P^2 \leq C(M^2 + 1/n)$.
+Otherwise, setting $K = \floor{M^2n}^{d/(2s + d)}$ balances squared bias and variance, and yields the claim.
+\qed.
+
+\paragraph{Testing.}
+In the goodness-of-fit testing problem, one asks for a test function---formally, a Borel measurable function $\phi$ that takes values in $\{0,1\}$--- which can distinguish between the hypotheses
+\begin{equation}
+\mathbf{H}_0: f_0 = f_0^{\star}, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{H}^{s}(\mc{X};M) \setminus \{f_0^{\star}\}.
+\end{equation} 
+To fix ideas, here and throughout we focus on the signal detection problem, meaning the special case where $f_0^{\star} = 0$.\footnote{This is without loss of generality since all the test statistics we consider are easily modified to handle the case when $f_0^{\ast}$ is not $0$, by simply subtracting $f_0^{\ast}(X_i)$ from each observation $Y_i$, with no change in the analysis.} For more background on nonparametric goodness-of-fit testing problems, see~\cite{ingster2012}.
+
+For the signal detection problem, the population-level spectral projection test $\wt{\varphi} := \1\{\wt{T} \geq K/N + \sqrt{2K/an^2}\}$ has bounded Type I error, $\Ebb_{0}[\wt{\varphi}] \leq a (1 + o(1))$.
+Proposition~\ref{prop:spectral_series_testing} gives an upper bound on the Type II error that holds over all $f_0 \in \mc{H}^s(\mc{X};M)$ for which $\|f_0\|_P^2$ is sufficiently large.
+\begin{proposition}
+	\label{prop:spectral_series_testing}
+	Suppose data is observed according to Model~\ref{def:model_flat_euclidean}, and that the density $p$ is known.
+Suppose additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$, $f_0 \in \mc{H}^{s}(\mc{X};M)$ for some $s > d/4$, and $\|f_0\|_{\Leb^4(\mc{X})}^4 \leq 1$.
+Then there exists a constant $C$ which does not depend on $f_0$, $M$ or $n$ such that the following statement holds: if the spectral projection test $\wt{\varphi}$ is computed with parameter $K = \max\{\floor{M^2n}^{2d/(4s + d)},1\}$, and if
+	\begin{equation}
+	\label{eqn:spectral_series_testing}
+	\|f_0\|_P^2 \geq C\min\Bigl\{M^2(M^2n)^{-4s/(4s + d)}, \frac{1}{n}\Bigr\}
+	\end{equation}
+	then the Type II error is upper bounded, $\Ebb_{f_0}[1 - \phi] \leq b$.
+\end{proposition}
+Assuming again that $n^{-1/2} \lesssim M$, the upper bound in~\eqref{eqn:spectral_series_testing} is $M^2(M^2n)^{-4s/(4s + d)}$, matching the usual minimax critical radius over Sobolev spaces (see e.g.~\cite{guerre02,ingster2009,ingster2012}.
+Specifically, \cite{ingster2009} show that the minimax squared critical radius is on the order of $n^{-4s/(4s + d)}$ when $M = 1$, and simple alterations of their analysis imply the rate $M^2(M^2n)^{-4s/(4s + d)}$ for general $M$.)
+
+On the other hand, when $s \leq d/4$ the minimax regression testing rates over $H^s(\mc{X})$ are not known.
+If one explicitly assumes $f_0 \in L^4(\mc{X};1)$---note that $H^s(\mc{X})$ does not continuously embed into $L^4(\mc{X})$ when $s \leq d/4$---then the minimax critical radius for regression testing is on the order of $n^{-1/2}$ \citep{guerre02}, and is achieved by a test using the naive statistic $\|{\bf Y}\|_n^2$.
+In other words, the regression testing problem over Sobolev spaces fundamentally changes when $s \leq d/4$, and hereafter when we discuss testing we will limit our consideration to $s > d/4$.
+\paragraph{Proof of Proposition~\ref{prop:spectral_series_testing}.}
+We briefly lay out the main ideas needed to prove Proposition~\ref{prop:spectral_series_testing}, following the lead of~\cite{ingster2009} who prove a similar result in the special case where $M = 1$ and $P$ is the uniform distribution over $\mc{X} = [0,1]^d$, and referring to that work for more details.
+
+We begin by computing the first two moments of the test statistic $\wt{T}$. The expectation is
+\begin{align*}
+\Ebb[\wt{T}] = \frac{(n - 1)}{n} \sum_{k = 1}^{K} \dotp{f_0}{\psi_K}_P^2 + \frac{K}{n} + \Ebb\Bigl[(f_0(X))^2 \dot \sum_{k = 1}^{K} (\psi_k(X))^2\Bigr],
+\end{align*}
+and from~\eqref{eqn:weyl} (Weyl's Law) we have that under the alternative $f_0 \neq 0$,
+\begin{equation*}
+\Ebb_{f_0}[\wt{T}] \geq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\lambda_{K + 1}^s(\Delta_{P})} + \frac{K}{n}.
+\end{equation*}
+To compute the variance, we decompose $\wt{T} = \wt{T}_{1,1} + \wt{T}_{1,2} + \wt{T}_{1,3} + \wt{T}_2$ into the sum of 3 U-statistics and the remaining diagonal terms, defined in terms of the equivalent kernel $\kappa(x,x') = \sum_{k = 1}^{K} \psi_k(x) \psi_k(x')$ as,
+\begin{align*}
+T_{1,1} & := \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} w_i w_j \kappa(X_i,X_j),\quad && T_{1,2} := \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} \bigl(w_i f_0(X_j) + w_jf_0(X_i)\bigr)\kappa(X_i,X_j) \\
+T_{1,3} & := \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} f_0(X_i) f_0(X_j) \kappa(X_i,X_j), \quad && T_2 := \frac{1}{n^2} \sum_{i = 1}^{n} Y_i^2 \kappa(X_i,X_i).
+\end{align*}
+The variances of each statistic can be found by routine computation (see~\cite{ingster2009}), and in particular satisfy the upper bounds
+\begin{align*}
+\Var(T_{1,1}) & \leq \frac{2K}{n^2}, \quad && \Var(T_{1,2}) \overset{\mathrm{(i)}}{\leq} \frac{C}{n}\|f_0\|_P^2\\
+\Var(T_{1,3}) & \overset{\mathrm{(ii)}}{\leq} C\biggl(\frac{K}{n}\|f_0\|_P^4 + \frac{K}{n^2}\|f_0\|_{L^4(\mc{X})}^4\biggr), \quad && \Var(T_{2}) \overset{\mathrm{(iii)}}{\leq} \frac{CK^2}{n^3}\biggl(1 + \|f_0\|_{L^4(\mc{X})}^4\biggr)
+\end{align*}
+where $\mathrm{(i)}-\mathrm{(iii)}$ hold due to local Weyl's law, i.e.~\eqref{eqn:local_weyl}. Upper bounds on Type I and Type II error,
+\begin{equation*}
+\Ebb_0[\wt{\varphi}] \leq \biggl(1 + CK/n^2\biggr)a, \quad \Ebb_{f_0}[1 - \wt{\varphi}] \leq \frac{C(K/n^2 + \|f_0\|_P^2 + K/n \|f_0\|_P^4 + K/n^2 \|f_0\|_{L^4(\mc{X})}^4)}{(\sum_{k = 1}^{K} \dotp{f_0}{\psi_k}_P^2 - \sqrt{2K/an})^2},
+\end{equation*}
+follow from Chebyshev's inequality. It can be verified that so long as
+\begin{equation}
+\label{pf:spectral_series_test}
+\|f_0\|_P^2 \geq C\Biggl(\frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\lambda_{K + 1}^s(\Delta_{P})} + \frac{\sqrt{K}}{n}\biggl(\sqrt{\frac{1}{a}} + \sqrt{\frac{1}{b}}\biggr)\Biggr)
+\end{equation}
+for a sufficiently large constant $C$, then $\Ebb_{f_0}[1 - \wt{\varphi}] \leq b$.
+The two summands in~\eqref{pf:spectral_series_test} are bias and standard deviation terms, respectively.
+When $M^2 \leq n^{-1}$, setting $K = 1$ gives the desired result.
+Otherwise, choosing $K = \floor{M^2n}^{2d/(4s + d)}$ balances these two terms, and leads to~\eqref{eqn:spectral_series_testing}.
+\qed
+% (AG 8/19/21): Same comment as above, regarding the role of M in random design regression testing.
+
+The main takeaway from Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} is that spectral projection methods for nonparametric regression achieve optimal rates of convergence, when the regression function $f_0$ is Sobolev smooth and the design density $p$ is known a priori and satisfies an appropriate notion of smoothness\footnote{The assumption $p \in C^{\infty}(\mc{X})$ could likely be weakened, but since this would not substantially add to the main points of Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, we do not pursue the details further.}.
+As we will see, Laplacian eigenmaps achieves comparable rates of convergence when $p$ is sufficiently smooth but potentially unknown, and in this sense both learns and leverages the design distribution in a way that more classical spectral projection methods do not.
+
+Of course, it is worth pointing out that other methods besides Laplacian eigenmaps are statistically optimal for nonparametric regression even when $p$ is unknown.
+We comment more on some of these in Section~\ref{sec:discussion}, after we have derived our major results regarding Laplacian eigenmaps.
 
@@
 
-% As previously explained, Laplacian eigenmaps is a discrete and noisy approximation to a spectral projection method using the eigenfunctions of $\Delta_P$.
-This is particularly useful when $P$ is unknown, or when the eigenfunctions of $\Delta_P$ cannot be explicitly computed.
-Our goal is to show that Laplacian eigenmaps methods are rate-optimal, notwithstanding the potential extra error incurred by this approximation.
-In this section and the following one, we will see that this is indeed the case: the estimator $\wh{f}$, and a test using the statistic $\wh{T}$, achieve optimal estimation and goodness-of-fit testing rates over Sobolev classes.
-% (AG 8/26/21): This segment is now redundant.
-
-In this section we give upper bounds on the error of Laplacian Eigenmaps in the flat Euclidean case, where we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to Model~\ref{def:model_flat_euclidean}.
-We will divide our theorem statements based on the regression function $f_0$ belongs to the first order Sobolev class $H^1(\mc{X})$ or a higher-order Sobolev class ($H_0^{s}(\mc{X})$ for some $s > 1$), since the details of the two settings are somewhat different.
+In this section we give upper bounds on the error of Laplacian eigenmaps in the flat Euclidean setting, where we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to Model~\ref{def:model_flat_euclidean}.
+We will divide our theorem statements based on whether the regression function $f_0$ belongs to the first order Sobolev class $H^1(\mc{X})$ or a higher-order Sobolev class ($H_0^{s}(\mc{X})$ for some $s > 1$), since the details of the two settings are somewhat different.
 
@@
 
-\paragraph{Estimation.} Laplacian eigenmaps depends on the kernel $\eta$ and two tuning parameters, the graph radius $\varepsilon$ and number of eigenvectors $K$. We now give some assumptions on each.
+\paragraph{Estimation.} Laplacian eigenmaps depends on the kernel $\eta$ and two tuning parameters, the graph radius $\varepsilon$ and number of eigenvectors $K$.
+We will need to make some assumptions on each.
 \begin{enumerate}[label=(K\arabic*)]
@@
 	\begin{equation*}
-	\int_{\Rd} \eta(\|z\|) \,dz = 1.
+	\int_{\Rd} \eta(\|z\|) \,dz = 1,
 	\end{equation*}
@@
 We comment on these assumptions after stating our first main theorem, regarding the estimation error of Laplacian eigenmaps.
+The proof of this theorem, along with the proofs of all subsequent results, can be found in the Appendix.
 \begin{theorem}
@@
 \end{theorem}
-From~\eqref{eqn:laplacian_eigenmaps_estimation_fo} it follows immediately that when $M \asymp 1$, then with constant probability $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2/(2 + d)}$, matching the standard minimax estimation rate over Sobolev classes.
+From~\eqref{eqn:laplacian_eigenmaps_estimation_fo} it follows immediately that when $n^{-1/2} \lesssim M \lesssim n^{1/d}$, then with constant probability $\|\wh{f} - f_0\|_n^2 \lesssim M^2(M^2n)^{-2/(2 + d)}$, matching the minimax estimation rate over Sobolev classes.
 
@@
 \begin{itemize}
-	\item \emph{Radius of the Sobolev ball.} If $M = M_n$ satisfies $n^{-1/2} \lesssim M_n \lesssim n^{1/d}$, then Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} implies that $\|\wh{f} - f_0\|_n^2 \lesssim M^2(M^2n)^{-2/(2 + d)}$ with constant probability.
-This matches the upper bound on $\Ebb\|\wt{f} - f_0\|_P^2$ given in~\eqref{eqn:spectral_series_estimation}, including in its dependence on $M$.
-It also matches the known minimax rate for some related problems, including estimation in the Gaussian sequence model~\citep{johnstone2011}, and estimation with a fixed design when the dimension $d = 1$~\citep{vandergeer2000}.
-
+	\item \emph{Radius of the Sobolev ball.} 
 	When $M = o(n^{-1/2})$ then computing Laplacian eigenmaps with $K = 1$ achieves the parametric rate $\|\wh{f} - f_0\|_n^2 \lesssim n^{-1}$, and the zero-estimator $\wh{f} = 0$ achieves the better rate $\|\wh{f} - f_0\|_n^2 \lesssim M^2$.
@@
 
-	\item \emph{In-sample error.} As we've already noted, since the Laplacian eigenmaps estimator is defined only at the design points $\{X_1,\ldots,X_n\}$, we will use the empirical norm $\|\cdot\|_n^2$ as our estimation loss, and return to the question of out-of-sample estimation later in Section~\ref{sec:out_of_sample}.
+	\item \emph{In-sample error.} Since the Laplacian eigenmaps estimator is defined only at the design points $\{X_1,\ldots,X_n\}$, we use the empirical norm $\|\cdot\|_n^2$ as our estimation loss.
+We return to the question of out-of-sample estimation later in Section~\ref{sec:out_of_sample}.
 
@@
 	In practice, one typically tunes hyper-parameters by cross-validation.
-However, because the estimator $\wh{f}$ is defined only in-sample, neither cross-validation nor any other sample-splitting method can be used to tune parameters for Laplacian eigenmaps.
+However, because the estimator $\wh{f}$ is defined only in-sample, neither cross-validation nor any other sample-splitting technique can be used to tune parameters for Laplacian eigenmaps.
 We return to this issue in Section~\ref{sec:out_of_sample}, when we propose an out-of-sample extension of $\wh{f}$.
@@
 The upper bound given in~\eqref{eqn:laplacian_eigenmaps_estimation_fo} holds with ``constant probability'', meaning with probability $1 - \delta - o(1)$.
-Under the stronger assumption that $f_0$ is $M$-Lipschitz, we can establish the same guarantee~\eqref{eqn:laplacian_eigenmaps_estimation_fo} with probability $1 - \delta^2/n - Cn\exp(-cn\varepsilon^d) - \exp(-K)$; in other words, we can give a high probability guarantee (for details see~\citep{green2021}).
+Under the stronger assumption that $f_0$ is $M$-Lipschitz, we can establish the same guarantee~\eqref{eqn:laplacian_eigenmaps_estimation_fo} with probability $1 - \delta^2/n - Cn\exp(-cn\varepsilon^d) - \exp(-K)$; in other words, we can give a high probability guarantee (for details see~\cite{green2021}).
 In this case a routine calculation shows that $\Ebb[\|\wh{f} - f_0\|_n^2]$ will also be on the some order as~\eqref{eqn:laplacian_eigenmaps_estimation_fo}.
@@
 Then $\mathbb{E}_0[\varphi] \leq a$, i.e $\varphi$ is a level-$a$ test.
-Suppose additionally $f_0 \in H^1(\mc{X},M)$, and that $d < 4$.
-Then there exist constants $C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_fo}, and if $f_0$ satisfies
+Suppose additionally $f_0 \in H^1(\mc{X};M)$, and that $d < 4$.
+Then there exist constants $C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n \geq N$: if the Laplacian eigenmaps test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_fo}, and if $f_0$ satisfies
 	\begin{equation}
@@
 Although~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} involves taking the maximum of several different terms, the important takeaway of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} is that if $n^{-1/2} \lesssim M \lesssim n^{(4 - d)/4d}$, then $\varphi$ has small worst-case risk as long as $f_0$ is separated from $0$ by at least $M^2(M^2n)^{-4/(4 + d)}$.
-In the special case $M \asymp 1$, this matches the minimax squared critical radius $n^{-4/(4 + d)}$, implying that $\varphi$ is a minimax rate-optimal test over $H^1(\mc{X};M)$ when $d \in \{1,2,3\}$.
+This implies that $\varphi$ is a minimax rate-optimal test over $H^1(\mc{X};M)$ when $d \in \{1,2,3\}$.
 As mentioned previously, when $d \geq 4$ the first order Sobolev space $H^1(\mc{X})$ does not continuously embed into $\Leb^4(\mc{X})$, and in this case the optimal rates for regression testing over Sobolev spaces are unknown.
@@
 \label{sec:higher_order_sobolev_classes}
-We now consider the situation where the regression function displays some higher-order regularity, $f_0 \in H_0^s(\mc{X})$.
+We now consider the situation where the regression function displays some higher-order regularity, $f_0 \in H_0^s(\mc{X};M)$.
 We show that Laplacian eigenmaps methods continue to be optimal for all orders of $s$, as long as the design density is itself also sufficiently regular, $p \in C^{s - 1}(\mc{X})$.
@@
 \end{enumerate}
-Crucially, when $n$ is sufficiently large the two conditions in~\ref{asmp:parameters_estimation_ho} are guaranteed to not be mutually exclusive.
-This is because so long as $M^2 = \omega(n^{-1})$ then $(M^2n)^{-2/(2(s - 1) + d)} = o((M^2n)^{-2/(2s + d)})$, regardless of $s$ and $d$.
+Crucially, when $n$ is sufficiently large the two conditions in~\ref{asmp:parameters_estimation_ho} are not mutually exclusive.
 \begin{theorem}
@@
 As compared to a sparse $G$, this results in more non-zero entries in the graph Laplacian, and increases the computational burden involved in computing $\wh{f}$.
-To address this issue, in Appendix~\ref{sec:computational_considerations} we review some approaches to \emph{spectral sparsification}, in which one efficiently computes a sparse graph $\wc{G}$ that approximates $G$ in a spectral sense (see).
-Letting $\wc{f}$ be the Laplacian eigenmaps estimator computed with respect to the sparsified graph $\wc{G}$, we provide upper bounds on $\|\wc{f} - f_0\|_n^2$, which show that under mild conditions on $\wc{G}$---which provably achieved by many techniques for spectral sparsification---the estimator $\wc{f}$ achieves the same rates of convergence as $\wh{f}$.
+To address this issue, in Appendix~\ref{sec:computational_considerations} we review some approaches to \emph{spectral sparsification}, in which one efficiently computes a sparse graph $\wc{G}$ that approximates $G$ in a spectral sense.
+The hope is that the Laplacian eigenmaps estimator $\wc{f}$ computed with respect to the sparsified graph $\wc{G}$ has similar statistical properties as $\wh{f}$, while being much faster to compute.
+To that end, we provide upper bounds on $\|\wc{f} - f_0\|_n^2$, which show that under mild conditions on $\wc{G}$---provably achieved by many spectral sparsification algorithms---the estimator $\wc{f}$ achieves the same rates of convergence as $\wh{f}$.
 \end{itemize}
@@
 \end{enumerate}
-When $d \leq 4$, for any value of $s \in \mathbb{N}$ when $n$ is sufficiently large it is possible to choose $\varepsilon$ and $K$ such that both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho} are satisfied, and our next theorem establishes that in this situation $\varphi$ is an optimal test.
+When $d \leq 4$ and $n$ is sufficiently large, it is possible to choose $\varepsilon$ and $K$ such that both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho} are satisfied, and our next theorem establishes that in this situation $\varphi$ is an optimal test.
 \begin{theorem}
@@
 \end{theorem}
-Similarly to the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that when $M \asymp 1$, then $\varphi$ is a minimax rate-optimal test over $H_0^s(\mc{X})$.
-However, unlike the first-order case, when $4 < d < 4s$ the minimax testing rate over $H_0^s(\mc{X})$ is still on the order of $M^2(M^2n)^{-4s/(4s + d)}$; unfortunately, we can no longer claim that $\varphi$ is an optimal test in this regime.
+Similarly to the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that when $n^{-1/2} \lesssim M \lesssim n^{(4s - d)/4d}$, then $\varphi$ is a minimax rate-optimal test over $H_0^s(\mc{X})$.
+However, unlike the first-order case, when $4 < d < 4s$ the minimax testing rate over $H_0^s(\mc{X})$ is still on the order of $M^2(M^2n)^{-4s/(4s + d)}$, but we can no longer claim that $\varphi$ is an optimal test in this regime.
 \begin{theorem}
@@
 \end{theorem}
-Focusing again on the special case where $M \asymp 1$, Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} says that $\varphi$ has have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2(s - 1) + d)}$ and $4 < d < 4s$.
+Focusing on the special case where $M \asymp 1$, Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} says that $\varphi$ has have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2(s - 1) + d)}$ and $4 < d < 4s$.
 This is smaller than the estimation rate $n^{-2s/(2s + d)}$, but larger than the minimax squared critical radius $n^{-4s/(4s + d)}$.
-As a technical matter, the problem is that when $d > 4$ there do not exist any choices of $\varepsilon$ and $K$ which satisfy both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho}, and as a result we cannot optimally balance (our upper bounds on) testing bias and variance (defined momentarily in~\eqref{eqn:testing_biasvariance}).
-Although we suspect $\varphi$ is truly suboptimal when $d > 4$, the inequality in~\eqref{eqn:testing_biasvariance} gives only an upper bound on testing bias, and thus we cannot rule out that the test $\varphi$ is optimal for all $4 < d < 4s$.
+
+% (AG 8/30/21): Is the below chunk helpful? Or not intuitive enough to be reasonable.
+It is intuitively reasonable that data-dependent spectral projection methods, such as Laplacian eigenmaps, should have more difficulty achieving the minimax rates of convergence for testing, as opposed to estimation.
+Optimally balancing the bias and variance of spectral projection tests (for instance~\eqref{pf:spectral_series_test}) requires choosing more and higher-frequency eigenvectors than in estimation.
+It is known~\citep{trillos2019} that higher-frequency graph Laplacian eigenvectors---which correspond to larger eigenvalues---are meaningfully worse approximations of their population-level limits.
+Put more simply, minimax optimal testing when $d > 4$ appears to require stronger approximation properties than graph Laplacian eigenvectors can satisfy.
+That being said, although we suspect $\varphi$ is truly suboptimal when $d > 4$, our analysis relies on an upper bound on testing bias.
+Since we do not prove a matching lower bound, we cannot rule out that the test $\varphi$ is optimal for all $s < d/4$.
 We leave the matter to future work.
@@
 We now outline the high-level strategy we follow when proving each of Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}.
-We analyze the estimation error of $\wh{f}$, and the testing error of $\wh{\varphi}$, by first conditioning on the design points $X_1,\ldots,X_n$ and deriving \emph{design-dependent} bias and variance terms.
+We analyze the estimation error of $\wh{f}$, and the testing error of $\varphi$, by first conditioning on the design points $\{X_1,\ldots,X_n\}$ and deriving \emph{design-dependent} bias and variance terms.
 For estimation, we have that with probability at least $1 - \exp(-K)$,
@@
 Comparing~\eqref{eqn:estimation_biasvariance} and~\eqref{eqn:testing_biasvariance} to~\eqref{pf:spectral_series_estimation_3} and~\eqref{pf:spectral_series_test}, we see that two continuum objects in the latter pair of bounds, the Sobolev norm $\|f_0\|_{\mc{H}^s(\mc{X})}^2$ and the eigenvalue $\lambda_k(\Delta_P)$, have been replaced by graph-based analogues: the graph Sobolev seminorm $\dotp{L_{n,\varepsilon}^sf_0}{f_0}_n$ and the graph Laplacian eigenvalue $\lambda_k^s$.
-These latter quantities, along with the empirical squared norm $\|f_0\|_n^2$, are random variables that depend on the random design points $X_1,\ldots,X_n$.
-We proceed to establish suitable upper and lower bounds that hold in probability.
-
-\paragraph{Estimates of graph quadratic forms.}
+These latter quantities, along with the empirical squared norm $\|f_0\|_n^2$, are random variables that depend on the random design points $\{X_1,\ldots,X_n\}$.
+We proceed to establish suitable upper and lower bounds on these quantities that hold in probability.
+
+\paragraph{Estimates of graph Sobolev seminorms.}
 In Proposition~\ref{prop:graph_seminorm_fo} we restate an upper bound on the Dirichlet energy $\dotp{L_{n,\varepsilon}f}{f}_n$ from~\cite{green2021}. 
@@
 \end{proposition}
-Proposition~\ref{prop:graph_seminorm_fo} follows by upper bounding the expectation of $\Ebb\dotp{L_{n,\varepsilon}f}{f}_n = \dotp{L_{P,\varepsilon}f}{f}_P$---where $L_{P,\varepsilon}$ is the non-local Laplacian operator defined in~\eqref{eqn:nonlocal_laplacian}---by (a constant times) the squared Sobolev norm $\|f\|_{H^1(\mc{X})}^2$, and an application of Markov's inequality.
-
-In this work, we establish that an analogous bound holds for $\dotp{L_{n,\varepsilon}^sf_0}{f_0}_n$ when $s > 1$. We call this quantity the \emph{graph Sobolev semi-norm}.
+Proposition~\ref{prop:graph_seminorm_fo} follows by upper bounding the expectation $\Ebb\dotp{L_{n,\varepsilon}f}{f}_n = \dotp{L_{P,\varepsilon}f}{f}_P$---where $L_{P,\varepsilon}$ is the non-local Laplacian operator defined in~\eqref{eqn:nonlocal_laplacian}---by (a constant times) the squared Sobolev norm $\|f\|_{H^1(\mc{X})}^2$, and then applying Markov's inequality.
+
+In this work, we establish that an analogous bound holds for the graph Sobolev seminorm $\dotp{L_{n,\varepsilon}^sf}{f}_n$, when $s > 1$.
 \begin{proposition}
@@
 \end{proposition}
-We now summarize the techniques used to prove Proposition~\ref{prop:graph_seminorm_ho}, which will help explain what role the conditions on $f_0$, $p$ and $\varepsilon$ play.
+We now summarize the techniques used to prove Proposition~\ref{prop:graph_seminorm_ho}, which will help explain the role played by our conditions on $f_0$, $p$ and $\varepsilon$.
 To upper bound $\dotp{L_{n,\varepsilon}^sf}{f}_n$ in terms of $\|f\|_{H^s(\mc{X})}^2$, we introduce an intermediate quantity: the \emph{non-local Sobolev seminorm} $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$.
@@
 Then the proof of Proposition~\ref{prop:graph_seminorm_ho} proceeds according to the following steps.
-\begin{itemize}
+\begin{enumerate}
 	\item First we note that $\dotp{L_{n,\varepsilon}^s f}{f}_n$ is itself a biased estimate of the non-local seminorm $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$.
@@
 Instead, we use the zero-trace property of $f$ to show that $L_{P,\varepsilon}^jf(x)$ is small.
-	\item Finally, we combine the results of previous two steps to deduce an upper bound on $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$ in terms of the squared Sobolev norm $\|f\|_{H^s(\mc{X})}^2$.
+	\item Finally, we combine the results of the previous two steps to deduce an upper bound on $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$ in terms of the squared Sobolev norm $\|f\|_{H^s(\mc{X})}^2$.
 Roughly speaking, when $s$ is odd, $\dotp{L_{P,\varepsilon}^sf}{f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}^jf) \approx \sigma_{\eta}^{2j}E_{P,\varepsilon}(\Delta_P^jf)$, whereas when $s$ is even $\dotp{L_{P,\varepsilon}^sf}{f}_P = \|L_{P,\varepsilon}L_{P,\varepsilon}^{j}f\|_{P}^2 \approx \sigma_{\eta}^{2j}\|L_{P,\varepsilon} \Delta_Pf\|_P^2$.
 Reasoning in this way, we can translate estimates of $L_{P,\varepsilon}^jf$ into an upper bound on the order-$s$ non-local Sobolev seminorm, even though $s > j$.
-\end{itemize}
+\end{enumerate}
 Together, these steps establish Proposition~\ref{prop:graph_seminorm_ho}.
@@
 They provide explicit bounds on the relative error $|\lambda_{k} - \lambda_{k}(\Delta_P)|/\lambda_{k}(\Delta_P)$, which show that the relative error is small for sufficiently large $n$ and small $\varepsilon$.
-Crucially, the guarantees hold simultaneously for all $1 \leq k \leq K$ as long as $\lambda_{K}(\Delta_P) = O(\varepsilon^{-2})$.
+Crucially, these guarantees hold simultaneously for all $1 \leq k \leq K$ as long as $\lambda_{K}(\Delta_P) = O(\varepsilon^{-2})$.
 These results are actually stronger than are necessary to establish Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho}---in order to get rate-optimality, we need only show that for the relevant values of $K$, $\lambda_{K}/\lambda_K(P) = \Omega_P(1)$---but unfortunately they all assume $P$ is supported on a manifold without boundary (i.e.
@@
 In contrast, at points $x$ near the boundary of $\mc{X}$, the graph Laplacian is known to approximate a different operator altogether \citep{belkin2012}.\footnote{This is directly related to the boundary bias of kernel smoothing, since the graph Laplacian can be viewed as a kernel-based estimator of $\Delta_P$.} This renders analysis of $\lambda_k$ substantially more challenging, since its continuum limit is not $\lambda_k(\Delta_P)$.
-Rather than establishing the convergence of $\lambda_k$, we will instead use Lemma~2 of \cite{green2021}, whose assumptions match our own, and who give a weaker bound on the ratio $\lambda_k/\lambda_k(\Delta_P)$ that will nevertheless suffice for our purposes.
+Rather than analyzing the convergence of $\lambda_k$, we will instead use Lemma~2 of \cite{green2021}, whose assumptions match our own, and who give a weaker bound on the ratio $\lambda_k/\lambda_k(\Delta_P)$ that will nevertheless suffice for our purposes.
 
@@
 Note carefully that the statement~\eqref{eqn:empirical_norm_sobolev} is \emph{not} a uniform guarantee over all $f \in H^s(\mc{X};M)$.
-Indeed, such a statement cannot hold in the sub-critical regime ($2s \leq d$).\footnote{This is because in the sub-critical regime, for any set of points $\{x_1,\ldots,x_n\}$ there exist a sequence of functions $(f_k)_{k \in \mathbb{N}} \subset H^s(\mc{X};1)$ satisfying $f_k(x_i) = 1$ for each $i = 1,\ldots,n$, but for which $\|f_k\|_P^2 \to 0$ as $k \to \infty$.} Fortunately, a pointwise bound---meaning a bound that holds with high probability for a single $f \in H^s(\mc{X})$---is sufficient for our purposes.
+Indeed, such a statement cannot hold in the sub-critical regime ($2s \leq d$).\footnote{This is because in the sub-critical regime, for any set of points $\{x_1,\ldots,x_n\}$ there exists a sequence of functions $\{f_k:k \in \mathbb{N}\} \subset H^s(\mc{X};1)$ satisfying $f_k(x_i) = 1$ for each $i = 1,\ldots,n$---and therefore $\|f_k\|_n^2 = 1$---but for which $\|f_k\|_P^2 \to 0$ as $k \to \infty$.} Fortunately, a pointwise bound---meaning a bound that holds with high probability for a single $f \in H^s(\mc{X})$---is sufficient for our purposes.
 
@@
 In this section we consider the manifold setting, where $(X_1,Y_1),\ldots,(X_n,Y_n)$ are observed according to Model~\ref{def:model_manifold}.
-In this setting, it is known that the minimax rates depend only on the intrinsic dimension $m$; more specifically,~\cite{bickel2007,ariascastro2018} show that for H\"{o}lder classes, the minimax estimation rate is $n^{-2s/(2s + m)}$ and the testing rate is $n^{-4s/(4s + m)}$\footnote{Although~\cite{ariascastro2018} considers density testing, usual arguments regarding equivalence of experiments~\cite{brown96} imply that the same rates apply to regression testing}.
-On the other hand, a theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
-We build on this work by showing that when $f_0 \in H^s(\mc{X})$ and $P$ is supported on a manifold, Laplacian eigenmaps achieve the sharper minimax estimation and testing rates.
-
-\subsection{Laplacian eigenmaps error rates under the manifold hypothesis}
-Unlike in the flat-Euclidean case, since Model~\ref{def:model_manifold} assumes that $\mc{X}$ is boundaryless it is easy to deal with the first-order $(s = 1)$ and higher-order $(s > 1)$ cases all at once.
+In this setting, it is known that the minimax rates depend only on the intrinsic dimension $m$; more specifically,~\cite{bickel2007,ariascastro2018} show that for functions with H\"{o}lder smoothness $s$, the minimax estimation rate is $n^{-2s/(2s + m)}$ and the testing rate is $n^{-4s/(4s + m)}$.\footnote{Although~\cite{ariascastro2018} considers density testing, usual arguments regarding equivalence of experiments~\citep{brown96} imply that the same rates apply to regression testing.} On the other hand, a theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+We build on this work by showing that when $P$ is supported on a manifold $\mc{X}$ and $f_0 \in H^s(\mc{X})$, Laplacian eigenmaps achieves the sharper minimax estimation and testing rates.
+
+\subsection{Upper bounds}
+Unlike in the flat-Euclidean case, since Model~\ref{def:model_manifold} assumes that $\mc{X}$ is without boundary it is easy to deal with the first-order $(s = 1)$ and higher-order $(s > 1)$ cases all at once.
 A more important distinction between the results of this section and those of Section~\ref{sec:minimax_optimal_laplacian_eigenmaps} is that we will establish Laplacian eigenmaps is optimal only when the regression function $f_0 \in H^s(\mc{X};M)$ for $s \in \{1,2,3\}$.
@@
 	\label{asmp:parameters_estimation_manifold}
-	For a constant $C_0$, the graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy
+	For constants $c_0,C_0$, the graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy
 	\begin{equation}
 	\label{eqn:radius_estimation_manifold}
-	C_0\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/m}, n^{-1/(2(s - 1) + m)}\biggr\} \leq \varepsilon \leq \min\{\mathrm{inj}(\mc{X}), K^{-1/m}\},
-	\end{equation}
-	where we recall that $\mathrm{inj}(\mc{X})$ is a lower bound on the injectivity radius of $\mc{X}$. Additionally,
+	C_0\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/m}, n^{-1/(2(s - 1) + m)}\biggr\} \leq \varepsilon \leq c_0\min\{1, K^{-1/m}\}.
+	\end{equation}
+	Additionally,
 	\begin{equation*}
@@
 \end{theorem}
-
-Unlike in the full-dimensional case, in the manifold setting our upper bounds on the estimation and testing error of Laplacian eigenmaps match the minimax rate only when $s \leq 3$.
-When $s \geq 4$, the containment $H^s(\mc{X};1) \subset H^{3}(\mc{X};1)$ (taking the radius of the Sobolev ball $M = 1$ for simplicity) implies that the Laplacian eigenmaps estimator $\wh{f}$ has in-sample mean-squared error of at most $n^{-6/(6 + m)}$, and that the Laplacian eigenmaps test has small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-12/(12 + d)}$; however, these are slower than the usual minimax rates.
+Focusing on the case $M \asymp 1$,\footnote{To the best of our knowledge, the minimax rates for general $M$ in the manifold setting have not been worked out.} the upper bounds in Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold} imply that Laplacian eigenmaps attain the optimal rates of convergence over Sobolev balls $H^s(\mc{X})$ for $s \in \{1,2,3\}$.
+
+Unlike in the full-dimensional case, in the manifold setting our upper bounds on the estimation and testing error of Laplacian eigenmaps do not match the minimax rate when $s \geq 4$.
+In this case, the containment $H^s(\mc{X};1) \subset H^{3}(\mc{X};1)$ implies that the Laplacian eigenmaps estimator $\wh{f}$ has in-sample mean-squared error of at most on the order of $n^{-6/(6 + m)}$, and that the Laplacian eigenmaps test has small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-12/(12 + m)}$; however, these are slower than the minimax rates.
 
@@
 However this is not an option for us, as we assume $\mc{X}$---and in particular its geodesics---are unknown.
-Likewise, a classical spectral projection estimator, using eigenfunctions of the manifold Laplace-Beltrami operator, will achieve the minimax rate for all values of $s$ and $m$; but this is undesirable for the same reason---we do not want to assume that $\mc{X}$ is known.
-It is not clear whether this gap between spectral projection and Laplacian eigenmaps estimators---or more generally, between estimators which assume the manifold is known, and those which do not---is real, or a product of loose upper bounds.
+Likewise, a spectral projection estimator using eigenfunctions of the manifold Laplace-Beltrami operator, will achieve the minimax rate for all values of $s$ and $m$; but this is undesirable for the same reason---we do not want to assume that $\mc{X}$ is known.
+It is not clear whether this gap between spectral projection and Laplacian eigenmaps estimators is real, or a product of loose upper bounds.
 
@@
 % (SB via AG): Maybe some comment on the difference between testing and estimation.
+% (AG 8/30/21): I put in a chunk after our higher-order testing theorems, commenting on the difference between testing and estimation.
 
@@
 \label{sec:out_of_sample}
-Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity} show that $\wh{f}$ is a minimax optimal estimator over Sobolev spaces, when the loss is  \emph{in-sample}---that is, measured in $\Leb^2(P_n)$ norm.
-In this section we consider \emph{out-of-sample} error---error measured in $L^2(P)$ norm---which is the more typical metric used in the study of nonparametric regression with random design.
-
-Of course, the Laplacian eigenmaps estimator is only defined at the design points $\{X_1,\ldots,X_n\}$, and to measure error in $L^2(P)$ norm we must first extend it to be defined over the entire domain $\Xset$.
+Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity} show that $\wh{f}$ is a minimax optimal estimator over Sobolev spaces, when the error is measured \emph{in-sample}---that is, in $\Leb^2(P_n)$ norm.
+In this section we consider \emph{out-of-sample} error---measured in $L^2(P)$ norm---which is the more typical metric used in the study of nonparametric regression with random design.
+
+Of course, the Laplacian eigenmaps estimator is only defined at the design points $\{X_1,\ldots,X_n\}$, and to measure error in $L^2(P)$ norm we must first extend it to the entire domain $\Xset$.
 We propose a simple method, kernel smoothing, to do the job.
@@
 	\label{asmp:ho_kernel}
-	The kernel function $\psi$ is an order-$s$ kernel, meaning it satisfies
-	\begin{equation*}
-	\int_{-\infty}^{\infty} \psi(|z|) \,dz = 1, \quad \int_{-\infty}^{\infty} z^j \psi(|z|) \,dz = 0 ~~\textrm{for}~j = 1,\ldots, s + d - 2, \quad \textrm{and}~ \int_{-\infty}^{\infty} z^{s + d - 1} \psi(|z|) \,dz < \infty.
+	The kernel function $\psi$ is an order-$s$ kernel, meaning it satisfies $\int_{-\infty}^{\infty} \psi(|z|) \,dz = 1$, and if $s \geq 2$,
+	\begin{equation*}
+	\int_{-\infty}^{\infty} z^j \psi(|z|) \,dz = 0 ~~\textrm{for}~j = 1,\ldots, s + d - 2, \quad \textrm{and}~ \int_{-\infty}^{\infty} z^{s + d - 1} \psi(|z|) \,dz < \infty. 
 	\end{equation*}
 \end{enumerate}
-% (SB via AG): Give an example of a kernel which satisfies these requirements.
+Kernels satisfying~\ref{asmp:ho_kernel} can be constructed using linear combinations of (truncated) polynomials in a standard way; for instance, the boxcar kernel $\psi(z) = (1/2)\cdot\1\{|z| \leq 1\}$ is an order-$2$ kernel, and $\psi(z) = (9/26 - 15z^2/52)\cdot\1\{|z| \leq 1\}$ is an order-$3$ kernel.
 
@@
 Consequently, if $\wc{f}$ converges at the minimax rate in squared $L^2(P_n)$-norm, then $T_{n,h}\wc{f}$ will converge at the minimax rate in squared $L^2(P)$-norm.
-It follows immediately from Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} (when $s = 1$) or Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} (when $s > 1$), that $T_{n,h}\wh{f}$ achieves the optimal rate of convergence in $L^2(P)$.
+It follows immediately from Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} (when $s = 1$) or Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} (when $s > 1$) that $T_{n,h}\wh{f}$ achieves the optimal rate of convergence in $L^2(P)$.
 
@@
 		\end{equation*}
-		\item If $f_0 \in H_0^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$ for some $s \in \mathbb{N}, s > 1$, and the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_ho}, and the out-of-sample extension $T_{n,h}\wh{f}$ is computed with bandwidth $h = n^{-1/(2(s - 1) + d)}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel} and~\ref{asmp:ho_kernel}, then
+		\item If $f_0 \in H_0^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$ for some $s \in \mathbb{N}, s \geq 2$, and the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_ho}, and the out-of-sample extension $T_{n,h}\wh{f}$ is computed with bandwidth $h = n^{-1/(2(s - 1) + d)}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel} and~\ref{asmp:ho_kernel}, then
 		\begin{equation*}
@@
 	\item \emph{Data-driven parameter tuning}.
-Since $T_{n,h}\wh{f}$ is defined out-of-sample, we can use sample splitting or cross validation methods to tune hyperparameters, which we could not do for the original estimator $\wh{f}$.
+Since $T_{n,h}\wh{f}$ is defined out-of-sample, we can use sample splitting or cross-validation methods to tune hyperparameters, which we could not do for the original estimator $\wh{f}$.
 For instance, we can (i) split the sample into two halves, (ii) use the first half to compute $T_{n,h}\wh{f}$ for various values of $\varepsilon$, $h$, and $K$, (iii) choose the optimal values of these three hyperparameters by minimizing error on the held out set.
@@
 We believe similar arguments should imply that $T_{n,h}\wh{f}$ is adaptive in this sense when $\varepsilon,h$ and $K$ are chosen by sample splitting, but the details should be worked out carefully.
+
+	% (AG 8/31/21): For now, I punted on proving formal results regarding sample-splitting. I can return later during a revision. 
 	\item \emph{Alternatives to extension by kernel smoothing}.
@@
 One potential drawback to kernel smoothing is that it can change the value of the estimate at data, meaning $\wh{f} \neq (\wc{f}(X_1),\ldots, \wc{f}(X_n))$.
-More sophisticated procedures based on Nystr\"{o}m extension of the eigenvectors $v_1,\ldots,v_K$~\cite{fowlkes2004} would inherit the generality of kernel smoothing while being genuine interpolators; however, their properties are substantially more difficult to analyze.
-
-	%Finally, the Nystr\"{o}m method extends  eigenvectors $v_k$ to functions $f_k: \mc{X} \to \Reals$ as follows,
-	%\begin{equation*}
-	%L_{n,\varepsilon}f_k = \lambda_k f_k~~\textrm{on $\mc{X}$, such that}~~f_k(X_i) = %v_k(X_i)~~\textrm{for $i = 1,\ldots,n$.}
-	%\end{equation*}
-	%Here $L_{n,\varepsilon}$ should be thought of as an operator acting on functions $f \in C(\mc{X})$ as follows,
-	%\begin{equation*}
-	%L_{n,\varepsilon}f(x) = \sum_{i = 1}^{n} (f(x) - f(X_i))\eta\Bigl(\frac{\|X_i - x\|}{\varepsilon}\Bigr)
-	%\end{equation*}
-	% The Nystr\"{o}m approach to extension makes intriguing (re-)use of the graph Laplacian $L_{n,\varepsilon}$ and its eigenvectors $v_k$.
-Similar methods have been analyzed in the context of kernel ridge regression (references), but little is known about its application to graph Laplacians.
+More sophisticated procedures based on Nystr\"{o}m extension of the eigenvectors $\{v_1,\ldots,v_K\}$~\citep{fowlkes2004} inherit the generality of kernel smoothing while resulting in genuine extrapolation; however, the statistical properties of these procedures are substantially more difficult to analyze.
 \end{itemize}
@@
 
-In this section we empirically demonstrate that Laplacian Eigenmaps is a reasonably good alternative to spectral projection, even when $n$ is only moderately large.
-In order to compare the two methods, in our experiments we stick to the simple case where $P$ is the uniform distribution over $\mc{X} = [0,1]^d$, and we have simple closed-form expressions for the eigenfunctions of $\Delta_P$.
-In general, it is not easy to analytically compute these eigenfunctions, which is part of the appeal of Laplacian Eigenmaps.
+In this section we empirically demonstrate that Laplacian eigenmaps is a reasonably good alternative to population-level spectral projection, even at moderately large sample sizes $n$.
+In order to compare the two approaches, in our experiments we stick to the simple case where the design distribution $P$ is the uniform distribution over $\mc{X} = [-1,1]^d$, and we have simple closed-form expressions for the eigenfunctions of $\Delta_P$.
+In general, it is not easy to analytically compute these eigenfunctions, which is part of the appeal of Laplacian eigenmaps.
 
@@
 In our first experiment, we compare the mean-squared error of the Laplacian eigenmaps estimator $\wh{f}$  to that of its population-level counterpart $\wt{f}$.
-We vary the sample size from $n = 1000$ to $n = 4000$; sample $n$ design points $X_1,\ldots,X_n$ from the uniform distribution on the cube $[-1,1]^d$; and sample responses $Y_i$ according to~\eqref{eqn:model} with regression function $f_0 = M/\lambda_K^{s/2} \cdot \psi_K$ for $K \asymp n^{d/(2s + d)}$ (the pre-factor $M/\lambda_K^{s/2}$ is chosen so that $|f_0|_{H^s(\mc{X})}^2 = M^2$).
+We vary the sample size from $n = 1000$ to $n = 4000$; sample $n$ design points $\{X_1,\ldots,X_n\}$ from the uniform distribution on the cube $[-1,1]^d$; and sample responses $Y_i$ according to~\eqref{eqn:model} with regression function $f_0 = M/\lambda_K^{s/2} \cdot \psi_K$ for $K \asymp n^{d/(2s + d)}$ (the pre-factor $M/\lambda_K^{s/2}$ is chosen so that $|f_0|_{H^s(\mc{X})}^2 = M^2$).
 In Figure~\ref{fig:fig1} we show the in-sample mean-squared error of the two estimators as a function of $n$, for different dimensions $d$ and order of smoothness $s$.
 We see that all estimators have mean-squared error converging to zero at roughly the minimax rate.
-While unsurprisingly population-level spectral projection has smaller error, generally speaking that the error of Laplacian Eigenmaps gets closer to that of population-level spectral projection as $n$ gets larger.
+While unsurprisingly population-level spectral projection has smaller error, generally speaking the error of Laplacian eigenmaps approaches that of population-level spectral projection as $n$ gets larger.
 
 We also compare the out-of-sample estimation error, over a held-out test set, of Laplacian eigenmaps plus kernel smoothing to the spectral projection estimator.
-The story is much the same, supporting our theoretical claim that the additional error incurred by kernel smoothing of Laplacian Eigenmaps is negligible.
+The story is much the same, supporting our theoretical claim that the additional error incurred by kernel smoothing of Laplacian eigenmaps is negligible.
 
@@
 In our second experiment, we compare the Laplacian eigenmaps test $\varphi$ against the  population-level spectral projection test $\wt{\varphi}$.
-The setup, in terms of $n$ and $P$, is the same as that of our first experiment.
-To empirically evaluate the \textcolor{red}{critical radius} $\epsilon_n(\phi,H^1(\mc{X};M))$ of a test $\phi$, we compute $\phi$ for each $f_0 \in \mc{F} \subset H^1(\mc{X};M)$, where $\mc{F}$ is a discrete subset of $H^1(\mc{X};M)$.
-For each of $b = 1,2,\ldots,100$, we compute $\epsilon_b$, the smallest value of $\epsilon$ such that $R_n(\phi,\mc{F},\epsilon_b) \geq b/100$.
-Then we take $\bar{\epsilon} = 1/100 \cdot \sum_{b = 1}^{100}\epsilon_b$ to be our empirical measure of worst-case risk.
-In Figure~\ref{fig:fig2}, we see that the critical radii of both Laplacian eigenmaps and spectral projection tests are quite close to each other, and converge at roughly the minimax rate.
+The setup is generally the same as that of our first experiment, but to get an empirical estimate of the critical radius the details are necessarily somewhat more complicated.
+First we take $\mc{F} = \{M/\lambda_k^{s/2} \psi_k\}_{k = 1}^{n}$ to be a discrete subset of $H^1(\mc{X};M)$.
+Then, for each $f_0 \in \mc{F}$, we run a given test $\phi$ (either the Laplacian eigenmaps test $\phi = \varphi$, or the population-level spectral projection test $\phi = \wt{\varphi}$.) and record whether it was a false negative or true positive.
+We repeat this process over $100$ replications, giving a Monte Carlo estimate of the type II error $E_{f_0}[1 - \phi]$ for each $f_0 \in \mc{F}$.
+Finally, we take the smallest value of $\|f_0\|_P^2$ such $E_{f_0}[1 - \phi] \leq b$ as our estimate of the critical radius of $\phi$.
+
+In Figure~\ref{fig:fig2}, we see that the estimated critical radii of both Laplacian eigenmaps and population-level spectral projection tests are quite close to each other, and converge at roughly the minimax rate.
 \begin{figure*}[tb]
@@
 \paragraph{Tuning parameters.}
-Our first two experiments demonstrate that in terms of statistical error, Laplacian eigenmaps methods are reasonable replacements for spectral projection methods.
+Our first two experiments demonstrate that Laplacian eigenmaps methods have comparable statistical performance to spectral projection methods.
 Laplacian eigenmaps depends on two tuning parameters, and in our final experiment we investigate the importance of both, focusing now on estimation.
@@
 On the one hand, empirically it seems that the optimal choice of bandwidth parameter $h$ is usually smaller than $\varepsilon$, as suggested by our theory.
-On the other hand, for Laplacian eigenmaps plus kernel smoothing we see that mean-squared error curves as a function of $K$ are often quite close to their minima even when we choose many more eigenvectors than is optimal for Laplacian eigenmaps or spectral projection.
-This is not reflected in our theory, where we require that $K$ be chosen in the same tight range as was required for Laplacian Eigenmaps to be optimal in-sample.
-However, it does make intuitive sense: extension by kernel smoothing further attenuates the noise, making the algorithm more forgiving to overfitting during the Laplacian Eigenmaps step.
+On the other hand, for Laplacian eigenmaps plus kernel smoothing we see that as a function of $K$, the mean-squared error curves are close to their minima even when we choose many more eigenvectors than is optimal for Laplacian eigenmaps.
+This is not reflected in our theory, which requires that $K$ be chosen in the same tight range as was required for Laplacian eigenmaps to be optimal in-sample.
+However, it does make intuitive sense: extension by kernel smoothing further attenuates the noise, making the algorithm more forgiving to overfitting during the Laplacian eigenmaps step.
 
@@
 	\includegraphics[width=.245\textwidth]{figures/tuning/sobolev/mse_by_radius_2d_1s.pdf}  
-	\caption{Mean squared error of Laplacian Eigenmaps (\textcolor{red}{red}), Laplacian Eigenmaps plus kernel smoothing (\textcolor{blue}{blue}), and spectral projection (\textcolor{green}{green}) as a function of tuning parameters.
+	\caption{Mean squared error of Laplacian eigenmaps (\textcolor{red}{red}), Laplacian eigenmaps plus kernel smoothing (\textcolor{blue}{blue}), and spectral projection (\textcolor{green}{green}) as a function of tuning parameters.
 Top row: the same regression function $f_0$ as used in Figure~\ref{fig:fig1}.
@@
 In each panel, all tuning parameters except the one being varied are set to their optimal values.
-For Laplacian Eigenmaps plus kernel smoothing, circular points and a solid line are used to denote the error as a function of the graph radius $\varepsilon$, whereas triangular points and a dashed line are used to denote the error as a function of the bandwidth $h$.}
+For Laplacian eigenmaps plus kernel smoothing, circular points and a solid line are used to denote the error as a function of the graph radius $\varepsilon$, whereas triangular points and a dashed line are used to denote the error as a function of the bandwidth $h$.}
 	\label{fig:fig3}
@@
 
-\input{sections/discussion}
-
-% AG: I think this following might just be garbage, but I will leave it in in case it piques Ryan or Siva's interest.
-
-% Proposition~\ref{prop:graph_seminorm_ho}, and its proof, show that the iterated graph Laplacian $L_{n,\varepsilon}^s$ does a reasonable job of approximating a higher-order differential operator.
-The conditions we require --- that $\varepsilon$ be sufficiently large, that $f$ be zero-trace, and that $p$ have regularity of order $s - 1$ --- indicate that the iterated graph Laplacian is an imperfect, if adequate, tool for this job.
-This is no fault of the graph Laplacian.
-Rather, as pointed out by \textcolor{red}{(Sadhanala et al.
-2017)} its reflects that graph Laplacians were designed for a very general circumstance in which the only notions of derivation are first-differentials, divergences, and compositions of the two.
-Indeed, viewed in this light it is a pleasant surprise that iterated graph Laplacians approximate higher-order differentials at all.
-When one assumes Euclidean structure, as we do in this paper, there exist many methods (e.g.
-local linear embeddings, Hessian local linear embedding, local tangent space alignment) which leverage this structure to approximate differential operators in a more sophisticated manner.
-However, thus far little theoretical investigation has been done into even the pointwise behavior of these approaches.%
+\section{Discussion}
+\label{sec:discussion}
+
+In this work, we have derived upper bounds on the rates of convergence for methods based on Laplacian eigenmaps, which imply that in various settings these methods are minimax rate-optimal over Sobolev classes.
+Importantly, these upper bounds hold under nonparametric conditions on the design density $p$, and allow for $p$ to be unknown and, potentially, supported on a low-dimensional manifold.
+Our results help explain the practical success of methods which leverage graph Laplacian eigenvectors for regression.
+They also distinguish Laplacian eigenmaps from more traditional spectral projection procedures, which rely on a density-dependent basis and thus require the density be known a priori.
+
+Of course, there do exist other methods for nonparametric regression which achieve optimal rates of convergence under similar (or indeed weaker) conditions on $p$.
+These include other graph-based approaches---graph Laplacian regularization---methods besides spectral projection---e.g.
+kernel smoothing, local polynomial regression, thin-plate splines---and continuum spectral projection methods which use the eigenfunctions of an operator defined independently of $p$.
+To be clear, we do not advocate Laplacian eigenmaps over these alternatives.
+Rather, we view our results as theoretically justifying a place for Laplacian eigenmaps in the nonparametric regression toolbox.
+
+That being said, Laplacian eigenmaps does have certain advantages over each of the aforementioned approaches.
+We now conclude by outlining some of these advantages (limiting our discussion to estimation):
+\begin{itemize}
+	\item \emph{Optimality over high-dimensional Sobolev spaces}.
+Graph Laplacian regularization achieves minimax optimal rates over $H^1(\mc{X})$ only when $d \in \{1,2,3,4\}$ \citep{sadhanala16, green2021}; when $d \geq 5$, graph Laplacian regularization appears to be suboptimal.
+In contrast, Laplacian eigenmaps is optimal over $H^1(\mc{X})$ for all dimensions $d$.
+An even more stark contrast can be drawn between Laplacian eigenmaps (and other spectral projection estimators) and the first order thin-plate spline, the solution of the variational problem
+	\begin{equation}
+	\underset{f \in H^1(\mc{X})}{\textrm{minimize}} \quad  \|{\bf Y} - f\|_n^2 + \lambda \|f\|_{H^1(\Reals^d)}^2,
+	\end{equation}
+	which is optimal and indeed well-posed only when $d = 1$.
+This appears to be related to a general phenomenon described by~\cite{dhillon2013,dicker2017}, in which principal components analysis followed by ordinary least-squares is shown to be always competitive with, and sometimes much better than, ridge regression.
+	\item \emph{Manifold adaptivity}.
+An oft-recommended alternative to the population-level spectral projection methods considered in Section~\ref{subsec:spectral_projection} is to run ordinary least-squares regression using eigenfunctions of a density-independent operator, such as the Laplacian $\Delta = \sum_{i = 1}^{d} \partial^2f/\partial x_i^2$.
+Under the conditions of Model~\ref{def:model_flat_euclidean}, such a method will indeed be rate-optimal, though the upper bounds may come with undesirably large constants if $p$ is very non-uniform.
+However, in contrast with Laplacian eigenmaps, under Model~\ref{def:model_manifold} this method cannot achieve the faster minimax rates of convergence.
+	\item \emph{Density adaptivity}.
+It is natural to ask whether the two stage estimator $T_{n,h}\wh{f}$ defined in Section~\ref{sec:out_of_sample} has any advantage over the simpler approach of directly kernel smoothing the responses, i.e.
+using the estimator $T_{n,h}{\bf Y}$ (possibly for a different choice of $h$).
+In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{n,h}\wh{f}\|_P^2$ is of a strictly lower order than $\inf_{h'} \Ebb\|f_0 - T_{n,h'}{\bf Y}\|_P^2$.
+This is possible because Laplacian eigenmaps induces a completely different bias than kernel smoothing.
+For example, when $f_0$ and $p$ satisfy the so-called \emph{cluster assumption}---meaning $f_0$ is piecewise constant in high-density regions (clusters) of $p$---then the bias of Laplacian eigenmaps can be much smaller than that of kernel smoothing (for equivalent levels of variance).
+
+	We emphasize that this does not contradict the well-known optimality properties of kernel smoothing over e.g.
+H\"{o}lder balls.
+Rather, in the standard nonparametric regression setup---which we adopt in the main part of this paper, and in which $P$ is assumed to be equivalent to Lebesgue measure---the biases of Laplacian eigenmaps and kernel smoothing happen to be equivalent.
+But when $P$ is sufficiently non-uniform, this is no longer the case.
+\end{itemize}
+Grounding each of these three points on a firmer and more complete theoretical basis would be, in our view, a valuable direction for future work.
 

### 2021-08-31 manuscripts/laplacian_eigenmaps/sections/discussion.tex
Made many small changes to Laplacian Eigenmaps paper. Also converted format back to something Siva would prefer.

@@
 
-In this work, we have derived upper bounds on the rates of convergence for methods based on Laplacian Eigenmaps, which imply that in various settings these methods are minimax rate-optimal over Sobolev classes.
+In this work, we have derived upper bounds on the rates of convergence for methods based on Laplacian eigenmaps, which imply that in various settings these methods are minimax rate-optimal over Sobolev classes.
 Importantly, these upper bounds hold under nonparametric conditions on the design density $p$, and allow for $p$ to be unknown and, potentially, supported on a low-dimensional manifold.
-Our results help explain the practical success of methods which leverage graph Laplacian eigenvectors in downstream learning problems.
-They also distinguish Laplacian Eigenmaps from more traditional spectral projection procedures, which rely on a density-dependent basis and thus require the density be known a priori.
+Our results help explain the practical success of methods which leverage graph Laplacian eigenvectors for regression.
+They also distinguish Laplacian eigenmaps from more traditional spectral projection procedures, which rely on a density-dependent basis and thus require the density be known a priori.
 
@@
 
-That being said, Laplacian eigenmaps can have advantages over each of the aforementioned approaches. We now conclude by outlining some of these advantages, limiting our discussion to estimation:
+That being said, Laplacian eigenmaps does have certain advantages over each of the aforementioned approaches.
+We now conclude by outlining some of these advantages (limiting our discussion to estimation):
 \begin{itemize}
-	\item \emph{Optimality over low-smoothness Sobolev spaces}.
+	\item \emph{Optimality over high-dimensional Sobolev spaces}.
 Graph Laplacian regularization achieves minimax optimal rates over $H^1(\mc{X})$ only when $d \in \{1,2,3,4\}$ \citep{sadhanala16, green2021}; when $d \geq 5$, graph Laplacian regularization appears to be suboptimal.
-In contrast, Laplacian eigenmaps is optimal for all dimensions $d$.
-An even more stark contrast can be drawn between Laplacian eigenmaps and the first order thin-plate spline, the solution of the variation problem
+In contrast, Laplacian eigenmaps is optimal over $H^1(\mc{X})$ for all dimensions $d$.
+An even more stark contrast can be drawn between Laplacian eigenmaps (and other spectral projection estimators) and the first order thin-plate spline, the solution of the variational problem
 	\begin{equation}
@@
 	\item \emph{Manifold adaptivity}.
-An oft-recommended alternative to the population-level spectral projection methods considered in Section~\ref{subsec:spectral_projection} is to run ordinary-least squares regression using eigenfunctions of a density-independent operator, such as the Laplacian operator $\Delta = \sum_{i = 1}^{d} \partial^2f/\partial x_i^2$.
+An oft-recommended alternative to the population-level spectral projection methods considered in Section~\ref{subsec:spectral_projection} is to run ordinary least-squares regression using eigenfunctions of a density-independent operator, such as the Laplacian $\Delta = \sum_{i = 1}^{d} \partial^2f/\partial x_i^2$.
 Under the conditions of Model~\ref{def:model_flat_euclidean}, such a method will indeed be rate-optimal, though the upper bounds may come with undesirably large constants if $p$ is very non-uniform.
@@
 using the estimator $T_{n,h}{\bf Y}$ (possibly for a different choice of $h$).
-In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{n,h}\wh{f}\|_P^2$ is of a strictly lower order than $\inf_{h'} \Ebb\|f_0 - T_{h',n}{\bf Y}\|_P^2$.
+In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{n,h}\wh{f}\|_P^2$ is of a strictly lower order than $\inf_{h'} \Ebb\|f_0 - T_{n,h'}{\bf Y}\|_P^2$.
 This is possible because Laplacian eigenmaps induces a completely different bias than kernel smoothing.
-For example, when $f_0$ and $p$ satisfy the so-called \emph{cluster} assumption--- e.g.
-$f_0$ is piecewise constant in high-density regions (clusters) of $p$--- then the bias of Laplacian eigenmaps can much smaller than that of kernel smoothing (for equivalent levels of variance).
+For example, when $f_0$ and $p$ satisfy the so-called \emph{cluster assumption}---meaning $f_0$ is piecewise constant in high-density regions (clusters) of $p$---then the bias of Laplacian eigenmaps can be much smaller than that of kernel smoothing (for equivalent levels of variance).
 
@@
 Rather, in the standard nonparametric regression setup---which we adopt in the main part of this paper, and in which $P$ is assumed to be equivalent to Lebesgue measure---the biases of Laplacian eigenmaps and kernel smoothing happen to be equivalent.
-But in general this is not the case.
+But when $P$ is sufficiently non-uniform, this is no longer the case.
 \end{itemize}

### 2021-08-31 manuscripts/laplacian_eigenmaps/sections/introduction.tex
Made many small changes to Laplacian Eigenmaps paper. Also converted format back to something Siva would prefer.

@@
 \label{sec:introduction}
-Two fundamental statistical goals are to learn and leverage structure of an unknown distribution, on the basis of samples drawn from said distribution.
+Two fundamental statistical goals are to learn and leverage structure of an unknown distribution on the basis of samples.
 Neighborhood graph Laplacians, and their associated spectra, have proven to be powerful and general tools for both purposes.
@@
 
-By now there exists a rich literature (e.g.
-\citep{koltchinskii2000,belkin07,vonluxburg2008,burago2014,shi2015,singer2017,garciatrillos18,trillos2019, calder2019, cheng2021,dunson2021}) explaining this practical success from a theoretical perspective.
-Loosely speaking, these works justify graph spectral methods by showing that the eigenvectors and eigenvalues of graph Laplacians are empirical approximations of population-level objects: the eigenvalues and eigenfunctions of a density-weighted Laplace-Beltrami operator.
+By now there exists a rich literature \citep{koltchinskii2000,belkin07,vonluxburg2008,burago2014,shi2015,singer2017,garciatrillos18,trillos2019, calder2019, cheng2021,dunson2021} explaining this practical success from a theoretical perspective.
+Loosely speaking, these works justify graph spectral methods by showing that the eigenvectors and eigenvalues of graph Laplacians are empirical approximations of population-level objects: the eigenvalues and eigenfunctions of a density-weighted Laplacian operator.
 These population-level objects in turn characterize various structural aspects of the density, such as its high- and low-density regions, the shape and intrinsic dimension of its support, and so forth.
@@
 In contrast, in this paper we develop theory showing that graph spectral methods can effectively \emph{leverage} learned structure to perform downstream (supervised) learning tasks.
-Our main goal is to show that spectral projection procedures--which use eigenvectors of a neighborhood graph Laplacian as features in a linear regression problem---are statistically optimal methods for two problems in nonparametric regression: estimation and goodness-of-fit testing.
+Our main goal is to show that spectral projection procedures---which use eigenvectors of a neighborhood graph Laplacian as features in linear regression---are statistically optimal methods for two problems in nonparametric regression: estimation and goodness-of-fit testing.
 
-The regression methods we consider are based on \emph{Laplacian eigenmaps}, first introduced by~\cite{belkin03a} \footnote{Technically speaking, as used by~\cite{belkin03a} ``Laplacian eigenmaps'' refers to a spectral embedding procedure.
+The regression methods we consider are based on \emph{Laplacian eigenmaps}, first introduced by~\cite{belkin03a}.\footnote{Technically speaking, as used by~\cite{belkin03a} ``Laplacian eigenmaps'' refers to a spectral embedding procedure.
 What we call Laplacian eigenmaps is a two-step regression method: the first step is this spectral embedding, and the second step computes the estimator $\wh{f}$ by running ordinary least squares in embedding space.
-Since in this work we only ever consider the embedding within the context of this two-step regression method, we take the liberty of referring to the entire procedure as Laplacian eigenmaps.}.
-Given samples $(X_1,Y_1),\ldots,(X_n,Y_n)$, Laplacian eigenmaps first forms a neighborhood graph, with vertices at the design points $X_1,\ldots,X_n$ and weighted edges between pairs of sufficiently close points $X_i$ and $X_j$.
-The Laplacian eigenmaps estimator $\wh{f}$ is then computed using the spectral decomposition of an (unweighted) Laplacian defined over this neighborhood graph, by projecting the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto a subspace spanned by a certain number of ``low-frequency'' graph Laplacian eigenvectors; more precisely, these are the eigenvectors corresponding to the smallest eigenvalues of the graph Laplacian.
+Since in this work we only ever consider the embedding within the context of this two-step regression method, we take the liberty of referring to the entire procedure as Laplacian eigenmaps.} Given samples $(X_1,Y_1),\ldots,(X_n,Y_n)$, Laplacian eigenmaps first forms a neighborhood graph, with vertices at the design points $\{X_1,\ldots,X_n\}$ and weighted edges between pairs of sufficiently close points $X_i$ and $X_j$.
+The Laplacian eigenmaps estimator $\wh{f}$ is then computed using the spectral decomposition of an (unweighted) Laplacian defined over this neighborhood graph, by projecting the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto a subspace spanned by a certain number of ``low-frequency'' graph Laplacian eigenvectors; more precisely, the eigenvectors corresponding to the smallest eigenvalues of the graph Laplacian.
 The Laplacian eigenmaps test statistic $\wh{T} = \|\wh{f}\|_n^2$ can in turn be used as a statistic to detect whether any signal is present, that is, whether the regression function $f_0(x) = \mathbb{E}[Y|X = x]$ is anywhere non-zero.
 (The precise definition of these methods is given in Section~\ref{subsec:laplacian_eigenmaps}.) We study these methods through the lens of a minimax analysis.
-Our major contribution is to provide upper bounds on the error of the Laplacian eigenmaps estimator and test which imply that they achieve the minimax optimal rate of convergence over natural smoothness classes.
+Our major contribution is to provide upper bounds on the error of the Laplacian eigenmaps estimator and test which imply that they achieve the minimax optimal rates of convergence over natural smoothness classes.
 
 More specifically, we consider the worst-case risk of Laplacian eigenmaps over normed balls in Sobolev spaces.
-Roughly speaking, for a domain $\mc{X} \subseteq \Rd$, the order-$s$ (Hilbert-)Sobolev space $H^s(\mc{X})$ consists of those functions which are $s$-times weakly differentiable, with derivatives bounded in $L^2(\mc{X})$ norm.
-Equivalently, Sobolev spaces can be defined in a spectral manner;  letting $\psi_1,\psi_2,\ldots$ be the eigenfunctions of a Laplace-Beltrami operator defined on $\mc{X}$, Sobolev spaces consist of those functions $f$ whose Fourier coefficients $\theta_k := \int_{\mc{X}} f \psi_k$ are bounded in a particular sequence norm (see Section~\ref{subsec:spectral_projection}).
-The minimax estimation and testing rates for Sobolev spaces are generally speaking well-known---when using mean-squared error as the loss, the estimation rate is $n^{-2s/(2s + d)}$, and the testing rate is $n^{-4s/(4s + d)}$---and are achieved by various estimators.
-However, since Sobolev spaces admit a spectral characterization in terms of $\psi_1,\psi_2,\ldots$, classical spectral projection methods\footnote{We use ``classical'' to refer to spectral projection methods using eigenfunctions of a (fixed) Laplacian or Laplace-Beltrami operator, as distinct from Laplacian eigenmaps which uses eigenvectors of a (data-dependent) graph Laplacian matrix.} using these eigenfunctions as features are a particularly natural choice for regression over Sobolev spaces, and are in a sense canonical estimators for these function classes.
+Roughly speaking, for a domain $\mc{X} \subseteq \Rd$, the order-$s$ (Hilbert-)Sobolev space $H^s(\mc{X})$ consists of those functions which are $s$-times weakly differentiable, with order-$1$ up through order-$s$ derivatives bounded in $L^2(\mc{X})$ norm.
+Equivalently, Sobolev spaces can be defined in a spectral manner;  letting $\psi_1,\psi_2,\ldots$ be the eigenfunctions of a Laplacian operator defined with respect to $\mc{X}$, Sobolev spaces consist of those functions $f$ whose Fourier coefficients $\theta_k := \int_{\mc{X}} f \psi_k$ are bounded in a particular sequence norm (see Section~\ref{subsec:spectral_projection}).
+The minimax estimation and testing rates for Sobolev spaces are generally speaking well-known---when using mean-squared error as the loss, the estimation rate is $n^{-2s/(2s + d)}$, and the testing rate is $n^{-4s/(4s + d)}$---and are achieved by various methods.
+However, since Sobolev spaces admit a spectral characterization in terms of $\psi_1,\psi_2,\ldots$, spectral projection methods that use these eigenfunctions as features are a particularly natural choice for regression over Sobolev spaces, and are in a sense canonical estimators for these function classes.
 % (AG 8/13/21): Last clause makes a strong statement that I personally believe, but may irritate certain readers. Keep or remove?
@@
 
-Laplacian eigenmaps serves as a data-dependent approximation to these classical spectral projection methods.
-This data dependency can be very useful.
-Computing classical spectral methods requires a priori knowledge of the design distribution, since they are defined with respect to the eigenfunctions of a Laplace-Beltrami operator which itself depends on this distribution.
-Such a priori information is often not available.
-On the other hand, Laplacian eigenmaps implicitly learns structure from the design points $\{X_1,\ldots,X_n\}$, and can thus be used even when the design distribution is unknown.
-Of course, this comes with a tradeoff.
+Laplacian eigenmaps serves as a data-dependent approximation to these population-level spectral projection methods.\footnote{We use the term ``population-level'' to emphasize that these methods use eigenfunctions which are the infinite-data limits of graph Laplacian eigenvectors.} This data dependency can be very useful.
+As the name suggests, population-level spectral methods require a priori knowledge of the design distribution, since they are defined with respect to the eigenfunctions of a Laplacian operator which itself depends on this distribution.
+For this reason, these methods are typically studied only under restrictive conditions on the design; for instance, that the design points are uniformly distributed over the unit cube.
+In many situations such restrictive conditions are not satisfied, and it may not be reasonable to assume prior knowledge of the design distribution.
+In contrast to population-level methods, Laplacian eigenmaps implicitly learns structure from the design points $\{X_1,\ldots,X_n\}$, and can thus be used even when the design distribution is unknown.
+Of course, this comes with a trade off.
 Intuitively, using an empirical approximation to the ``right'' population-level basis will incur some additional error.
-The question is whether, notwithstanding this additional error, Laplacian eigenmaps shares the strong statistical properties of classical spectral projection.
+The question is whether, notwithstanding this additional error, Laplacian eigenmaps shares the strong statistical properties of its population-level counterpart.
 
@@
 \begin{itemize}
-	\item Over the Sobolev space $H^{s}(\mc{X})$, we establish the Laplacian eigenmaps estimator $\wh{f}$ has in-sample mean-square error on the order of $n^{-2s/(2s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d$.
+	\item Over the Sobolev space $H^{s}(\mc{X})$, we establish that the Laplacian eigenmaps estimator $\wh{f}$ has in-sample mean-squared error of at most on the order of $n^{-2s/(2s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d$.
 A test based on the statistic $\|\wh{f}\|_n^2$ has a squared critical radius on the order of $n^{-4s/(4s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d \in \{1,2,3,4\}$.
+	\item We also consider the behavior of Laplacian eigenmaps when the data satisfies a \emph{manifold hypothesis}, meaning the design points $\{X_1,\ldots,X_n\}$ belong to some low-dimensional submanifold of $\Rd$.
+In this case, it is known that the minimax rates over Sobolev spaces depend on the intrinsic dimension $m$ of the manifold: explicitly, the estimation rate is $n^{-2s/(2s + m)}$ and the testing rate is $n^{-4s/(4s + m)}$.
+We give upper bounds showing that Laplacian eigenmaps achieves these rates so long as $s \in \{1,2,3\}$.
+One can interpret these results as precisely quantifying a way in which Laplacian eigenmaps leverages structure, by adapting to the intrinsic dimension of the design.
 	\item The estimator $\wh{f}$ is defined only in-sample.
 We propose a simple out-of-sample extension based on kernel smoothing, and show that this extension adds negligible error, resulting in an estimator which is defined on all of $\mc{X}$ and has out-of-sample mean-squared error also on the order of $n^{-2s/(2s + d)}$.
-	\item We also consider the behavior of Laplacian eigenmaps when the data satisfies a \emph{manifold hypothesis}, meaning the design points $X_1,\ldots,X_n$ belong to some low-dimensional submanifold of $\Rd$.
-In this case, it is known that the minimax rates over Sobolev spaces depend on the intrinsic dimension $m$ of the manifold: explicitly, the estimation rate is $n^{-2s/(2s + m)}$ and the testing rate is $n^{-4s/(4s + m)}$.
-We give upper bounds implying that Laplacian eigenmaps achieves these rates so long as $s \in \{1,2,3\}$.
-One can interpret these results as precisely quantifying a way in which Laplacian Eigenmaps leverages structure, by adapting to the intrinsic dimension of the design.
 \end{itemize}
-% (AG 8/27/21): Removed the following sentence, because I no longer know if it is true.
-% In all cases, our bounds also depend optimally on the radius $M$ of the Sobolev ball under consideration.
+In all cases, our bounds also depend optimally on the radius $M$ of the Sobolev ball under consideration.
 
-To derive these results, we develop an framework for analysis reminiscent of that  classically used to analyze population-level spectral projection estimators, but ultimately quite different.
+To derive these results, we develop a framework for analysis reminiscent of that  classically used to analyze population-level spectral projection estimators, but ultimately quite different.
 The chief difference is in the approximation error (bias).
-For our purposes we must establish that with high probability, the eigenvectors of graph Laplacians (which are data-dependent and hence random objects) effectively approximate continuum Sobolev functions, whereas in the classical analysis an upper bound on the approximation error follows immediately from the aforementioned spectral characterization of Sobolev spaces.
-To tightly upper bound the approximation error of Laplacian eigenmaps, we rely on several recent results regarding concentration of the spectra of graph Laplacians around their population limits, and prove some additional results which may be of independent interest.
+For our purposes we must establish that with high probability, the eigenvectors of graph Laplacians (which are data-dependent and hence random objects) can be used to effectively approximate continuum Sobolev functions, whereas in the classical analysis an upper bound on the approximation error follows immediately from the aforementioned spectral characterization of Sobolev spaces.
+To tightly upper bound the approximation error of Laplacian eigenmaps, we rely on several recent results regarding concentration of the spectra of graph Laplacians around their population limits, and prove some additional results of this type which may be of independent interest.
 
-Our work establishes that Laplacian eigenmaps shares many of the optimality properties of classical spectral projection.
-However, for some values of $s$ (number of derivatives) and $d$ (dimension), there do exist gaps between our upper bounds on the error of Laplacian eigenmaps and the minimax rates.
+For some values of $s$ (number of derivatives) and $d$ (dimension), there do exist gaps between our upper bounds on the error of Laplacian eigenmaps and the minimax rates.
 Although we do not give corresponding lower bounds verifying the tightness of our analysis, we believe these gaps reflect the true behavior of the method rather than some looseness in our analysis, and we comment more on this at relevant parts in the text.
@@
 
-The majority of work on supervised learning using graphs adopts a \emph{fixed design} perspective, treating the design points $x_1,\ldots,x_n$ as vertices of a fixed graph, and carrying out inference with respect to the conditional mean vector $(f_0(x_1),\ldots,f_0(x_n))$.
-In this setting, tight upper bounds have been established that certify the optimality of graph-based methods for estimation \citep{wang2016,hutter2016,sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}) and testing \citep{sharpnack2010identifying,sharpnack2013b,sharpnack2013,sharpnack2015} over different ``function'' classes (in quotes because these classes really model the $n$-dimensional vector of evaluations).
-In particular,~\citet{sadhanala16} and~\citet{sharpnack2015} analyze the Laplacian eigenmaps estimator and test statistic, respectively.
+Most work on supervised learning using graphs adopts a \emph{fixed design} perspective, treating the design points $X_1 = x_1,\ldots,X_n = x_n$ as vertices of a fixed graph, and carrying out inference with respect to the conditional mean vector $(f_0(x_1),\ldots,f_0(x_n))$.
+In this setting, matching upper and lower bounds have been established that certify the optimality of graph-based methods for estimation \citep{wang2016,hutter2016,sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}) and testing \citep{sharpnack2010identifying,sharpnack2013b,sharpnack2013,sharpnack2015} over different ``function'' classes (in quotes because these classes really model the $n$-dimensional vector of evaluations).
 This setting is quite general, because the graph need not be a geometric graph defined on a vertex set which belongs to Euclidean space.
@@
 In the random design setting the analysis of graph-based approaches to supervised learning is more challenging, and only a few works have studied the matter.
-\cite{zhou2011} provide an asymptotic analysis of the Laplacian eigenmaps estimator in the semi-supervised setting, as the number of unlabeled samples tends to infinity; in this case the estimator reduces to the classical spectral projection estimator, and there is no extra error incurred by using data-dependent eigenvectors.
+\cite{zhou2011} provide an asymptotic analysis of the Laplacian eigenmaps estimator in the semi-supervised setting, as the number of unlabeled samples tends to infinity; in this case the estimator is equivalent to population-level spectral projection, and there is no extra error incurred by using data-dependent eigenvectors.
 \cite{lee2016} consider the \emph{diffusion maps} estimator---which uses the eigenvectors of a differently normalized graph Laplacian---in a similar supervised setting to our own, but prove suboptimal rates of convergence.
@@
 Laplacian regularization can, like Laplacian eigenmaps, be cast as a spectral method.
-The difference is that Laplacian regularization induces some smooth shrinkage based on the eigenvalue, where Laplacian eigenmaps corresponds to a hard spectral cutoff.
+The difference is that Laplacian regularization induces some smooth shrinkage based on the eigenvalue, whereas Laplacian eigenmaps corresponds to a hard spectral cutoff.
 Interestingly, this difference leads to a stark contrast in the ultimate statistical properties of the two approaches.
@@
 We now outline the structure of the rest of this paper.
-In Section~\ref{sec:setup_main_results}, we give our formal modeling assumptions, and precisely define the Laplacian Eigenmaps methods we study.
-Propositions~\ref{prop:spectral_series_estimation} and ~\ref{prop:spectral_series_testing}, in Section~\ref{subsec:spectral_projection}, show that under rather general (nonparametric) conditions on the design distribution, classical spectral projection methods achieve minimax rates of convergence over Sobolev classes.
-Then in Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps}, \ref{sec:out_of_sample} and and~\ref{sec:manifold_adaptivity}, we give our main upper bounds on the error of Laplacian eigenmaps (summarized above); these upper bounds hold under similarly general conditions, and imply that Laplacian Eigenmaps estimators and tests are also minimax rate-optimal.
-In Section~\ref{sec:experiments} we examine the empirical behavior of Laplacian eigenmaps, and show that even at moderate sample sizes Laplacian eigenmaps is competitive with classical spectral projection.
+In Section~\ref{sec:setup_main_results}, we give our formal modeling assumptions, and precisely define the Laplacian eigenmaps methods we study.
+Propositions~\ref{prop:spectral_series_estimation} and ~\ref{prop:spectral_series_testing}, in Section~\ref{subsec:spectral_projection}, show that under rather general (nonparametric) conditions on the design distribution, population-level spectral projection methods achieve minimax rates of convergence over Sobolev classes.
+Then in Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps},~\ref{sec:manifold_adaptivity}, and \ref{sec:out_of_sample}, we give our main upper bounds on the error of Laplacian eigenmaps (summarized above); these upper bounds hold under similarly general conditions, and imply that Laplacian eigenmaps estimators and tests are also minimax rate-optimal.
+In Section~\ref{sec:experiments} we examine the empirical behavior of Laplacian eigenmaps, and show that even at moderate sample sizes Laplacian eigenmaps is competitive with population-level spectral projection.
 We conclude with some discussion in Section~\ref{sec:discussion}.
@@
 We define $\dotp{f}{g}_P := \int fg\,dP$, and let $L^2(P)$ contain those functions $f$ for which $\|f\|_P^2 := \dotp{f}{f}_P$ is finite.
-Finally, we let $L^2(P_n)$ consist of those ``functions'' $f: \set{X_1,\ldots,X_n} \to \Reals$ for which the empirical norm $\|f\|_{n}^2 := \frac{1}{n}\sum_{i = 1}^{n} \bigl(f(X_i)\bigr)^2 < \infty$.
+Equivalently, we let $L^2(P_n)$ consist of those ``functions'' $f: \set{X_1,\ldots,X_n} \to \Reals$ for which the empirical norm $\|f\|_{n}^2 := \frac{1}{n}\sum_{i = 1}^{n} \bigl(f(X_i)\bigr)^2 < \infty$.
 When there is no chance of confusion, we will sometimes associate functions in $L^2(P_n)$ with vectors in $\Reals^n$, and vice versa.
 We use $C^k(\mc{X})$ to refer to functions which are $k$ times continuously differentiable in $\mc{X}$, either for some integer $k \geq 1$ or for $k = \infty$.
-We let $C_c^{\infty}(\mc{X})$ represent those functions in $C^{\infty}(\mc{X})$ with support $V$ compactly contained in $\mc{X}$, meaning $\wb{V} \subseteq \mc{X}$.
+We let $C_c^{\infty}(\mc{X})$ represent those functions in $C^{\infty}(\mc{X})$ with support $V$ compactly contained in $\mc{X}$, meaning $\wb{V}$ is compact and $\wb{V} \subseteq \mc{X}$.
 We write $\partial f/\partial r_i$ for the partial derivative of $f$ in the $i$th standard coordinate of $\Rd$, and use the multi-index notation $D^{\alpha}f := \partial^{|\alpha|}f/\partial^{\alpha_1}x_1\ldots\partial^{\alpha_d}x_d$ for multi-indices $\alpha \in \Reals^d$.

### 2021-08-31 manuscripts/laplacian_eigenmaps/sections/preliminaries.tex
Made many small changes to Laplacian Eigenmaps paper. Also converted format back to something Siva would prefer.

@@
 
-We begin in Sections~\ref{subsec:regression_laplacian_eigenmaps}-\ref{subsec:laplacian_eigenmaps} by precisely defining the models (random design points, Sobolev-smooth regression functions) and methods (Laplacian eigenmaps) under consideration.
-Then in Section~\ref{subsec:spectral_projection}, we connect Sobolev spaces to Laplacian eigenmaps using the spectrum of a density-dependent Laplace-Beltrami operator, and show in Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} that projection methods which use the eigenfunctions of this operator are statistically optimal.
+We begin in Sections~\ref{subsec:regression_laplacian_eigenmaps}-\ref{subsec:laplacian_eigenmaps} by precisely defining the models (random design points, Sobolev regression functions) and methods (Laplacian eigenmaps) under consideration.
+Then in Section~\ref{subsec:spectral_projection}, we connect Sobolev spaces to Laplacian eigenmaps using the spectrum of a density-dependent Laplacian operator, and show in Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} that projection methods which use the eigenfunctions of this operator are statistically optimal.
 
@@
 \end{equation}
-with regression function $f_0: \mc{X} \to \Reals$, and $w_i \sim N(0,1)$ independent Gaussian noise.)
+with regression function $f_0: \mc{X} \to \Reals$, and $w_i \sim N(0,1)$ independent Gaussian noise.
 % (AG 8/4/21): Removed the comment regarding noise with variance = \sigma^2.
@@
 % (AG 8/13/21): I got rid of the sentence ``The data are sampled according to (1).'' It seemed repetitive. 
-At various points we will also assume that the density $p \in C^k(\mc{X})$; when we make such assumptions we will state them explicitly.
-
-We model the regression function as belonging to an order-$s$ Sobolev space, and being bounded in Sobolev norm.
+At various points we will also assume that the density $p \in C^k(\mc{X})$.
+On the other hand, we model the regression function as belonging to an order-$s$ Sobolev space, and being bounded in Sobolev norm.
 \begin{definition}[Sobolev space on a flat Euclidean domain]
 	\label{def:sobolev_space}
-	For an integer $s \geq 1$, a function $f \in L^2(\mc{X})$ belongs to the Sobolev space $H^s(\mc{X})$ if for all $|\alpha| \leq s$, the weak derivatives $D^{\alpha}f$ exists and satisfy $D^{\alpha}f \in L^2(\mc{X})$.
+	For an integer $s \geq 1$, a function $f \in L^2(\mc{X})$ belongs to the Sobolev space $H^s(\mc{X})$ if for all $|\alpha| \leq s$, the weak derivatives $D^{\alpha}f$ exist and satisfy $D^{\alpha}f \in L^2(\mc{X})$.
 The $j$th order semi-norm for $f \in H^s(\mc{X})$ is $|f|_{H^j(\mc{X})} := \sum_{|\alpha| = j}\|D^{\alpha}f\|_{\Leb^2(\mc{X})}$, and the corresponding norm
@@
 \end{definition}
-Finally when $s > 1$ we will assume that $f_0$ satisfies a zero-trace boundary condition.
+When $s > 1$ we will also assume that $f_0$ satisfies a zero-trace boundary condition.
 Recall that $H^s(\mc{X})$ can alternatively be defined as the completion of $C^{\infty}(\mc{X})$ in the Sobolev norm $\|\cdot\|_{H^s(\mc{X})}$.
@@
 \paragraph{Manifold model.}
-As in the flat Euclidean case, we start with some regularity conditions on the design.
+As in the flat Euclidean case, we start with some regularity conditions on the design. One such condition will be on the reach $R$ of the manifold $\mc{X}$, which we recall is defined as follows:
+\begin{equation*}
+R := \Bigl\{\sup_{r > 0}: \forall z \in \Rd, \inf_{x \in \mc{X}} \|z - x\| \leq r, ~\exists! y \in \mc{X}~\mathrm{s.t.}~\|z - y\| = \inf_{x \in \mc{X}} \|z - x\|\Bigr\}.
+\end{equation*}
+In words, the reach is the largest radius of a ball which can be rolled around the manifold $\mc{X}$.
 \begin{definition}[Manifold model]
 	\label{def:model_manifold}
-	The support $\mc{X}$ of the design distribution $P$ is a closed, connected, and smooth Riemannian manifold (without boundary) embedded in $\Rd$, of intrinsic dimension $1 \leq m < d$, and with a positive injectivity radius $\mathrm{inj}(\mc{X}) > 0$.
+	The support $\mc{X}$ of the design distribution $P$ is a closed, connected, and smooth Riemannian manifold (without boundary) embedded in $\Rd$, of intrinsic dimension $1 \leq m < d$, and with a positive reach $R > 0$.
 The design distribution $P$ admits a Lipschitz density $p$ with respect to the volume form $d\mu$ induced by the Riemannian structure of $\mc{X}$, which is bounded away from $0$ and $\infty$,
@@
 \end{definition}
-In the above, recall that the injectivity radius $\mathrm{inj}(\mc{X})$ of an $m$-dimensional Riemannian manifold $\mc{X}$ is the maximum $\delta > 0$ such that the exponential map $\exp_x: B_m(0,\delta) \subset T_x(\mc{X}) \to B_{\mc{X}}(x,\delta) \subset \mc{X}$ is a diffeomorphism for all $x \in \mc{X}$.
 
@@
 Letting $r_1,\ldots,r_m$ be the standard basis of $\Reals^m$, for a given chart $(\phi,U)$ (meaning an open set $U \subseteq \mc{X}$, and a smooth mapping $\phi: U \to \Reals^m$) we write $\phi =: (x_1,\ldots,x_m)$ in local coordinates, meaning $x_i = r_i \circ \phi$.
-Then the partial derivative $\partial f/\partial x_i$ of a function $f$ with respect to $x_i$ at $x \in U$ is
+Then we define the partial derivative $\partial f/\partial x_i$ of a function $f: \mc{X} \to \Reals$ at $x \in U$ to be
 \begin{equation*}
@@
 \end{definition}
-The partial derivatives $D^{\alpha}f$ will depend on the choice of local coordinates, and so will the resulting Sobolev norm~$\|f\|_{H^s(\mc{X})}$.
-However, for our purposes the important point is that regardless of the choice of local coordinates the resulting norms will be equivalent\footnote{Recall that norms $\|\cdot\|_1$ and $\|\cdot\|_2$ on a space $\mc{F}$ are said to be equivalent if there exist constants $c$ and $C$ such that
+The partial derivatives $D^{\alpha}f$ clearly depend on the choice of local coordinates, and so will the resulting Sobolev norm~$\|f\|_{H^s(\mc{X})}$.
+However, for our purposes the important thing is that regardless of the choice of local coordinates the resulting norms will be equivalent\footnote{Recall that norms $\|\cdot\|_1$ and $\|\cdot\|_2$ on a space $\mc{F}$ are said to be equivalent if there exist constants $c$ and $C$ such that
 	\begin{equation*}
@@
 	\end{equation*}} 
-and so the ultimate Sobolev space $H^s(\mc{X})$ is independent of the choice of local coordinates.
+and so the ultimate Sobolev space $H^s(\mc{X})$ is independent of local coordinates.
 For more information regarding manifolds and Sobolev spaces defined thereupon, see~\cite{lee2013} and~\cite{hebey1996}.
 
-\subsection{Laplacian eigenmaps}
+\subsection{Laplacian Eigenmaps}
 \label{subsec:laplacian_eigenmaps}
@@
 The pre-factor $(n\varepsilon^{2 + \mathrm{dim}(\mc{X})})^{-1}$ ensures non-degenerate stable limits as $n \to \infty, \varepsilon \to 0$).
-Written in standard coordinates we have $(n\varepsilon^{\dim(\mc{X}) + 2}) \cdot L_{n,\varepsilon} = D - W$, where $D \in \Reals^{n \times n}$ is the diagonal degree matrix, $D_{ii} = \sum_{i = 1}^{n} W_{ij}$.
+Note that $(n\varepsilon^{\dim(\mc{X}) + 2}) \cdot L_{n,\varepsilon} = D - W$, where $D \in \Reals^{n \times n}$ is the diagonal degree matrix, $D_{ii} = \sum_{i = 1}^{n} W_{ij}$.
 
-The graph Laplacian is a positive semi-definite matrix, and admits the eigendecomposition $L_{n,\varepsilon} = \sum_{k = 1}^{n} \lambda_k v_k v_k^{\top}$, where for each $k = 1,\ldots,n$ the eigenvalue-eigenvector pair $(\lambda_k,v_k)$ satisfies
+The graph Laplacian is a positive semi-definite matrix, and admits the eigendecomposition $L_{n,\varepsilon} = \sum_{k = 1}^{n} \lambda_k v_k v_k^{\top}$, where for each $k \in \{1,\ldots,n\}$ the eigenvalue-eigenvector pair $(\lambda_k,v_k)$ satisfies
 \begin{equation*}
@@
 \end{equation} 
-If $\wh{f}$ is a reasonable estimate of $f_0$, then the Laplacian eigenmaps test statistic
+If $\wh{f}$ is a reasonable estimate of $f_0$, then the Laplacian eigenmaps test statistic,
 \begin{equation}
 \label{eqn:laplacian_eigenmaps_test}
-\wh{T} := \|\wh{f}\|_n^2 = \frac{1}{n} {\bf Y}^{\top} V_K V_K^{\top} {\bf Y}
+\wh{T} := \|\wh{f}\|_n^2 = \frac{1}{n} {\bf Y}^{\top} V_K V_K^{\top} {\bf Y},
 \end{equation}
@@
 
-% AG 8/12/21: Commented out due to overlap with footnote in new introduction.
-% It may be helpful to comment briefly on the term ``Laplacian eigenmaps'', which we use a bit differently than is typical in the literature.
-Laplacian eigenmaps typically refers to an algorithm for embedding, which maps each design point $X_1,\ldots,X_n$ to $\Reals^K$ according to $X_i \mapsto (v_{1,i}, \ldots, v_{K,i})$.
-Viewing this embedding as a feature map, we can then interpret the estimator $\wh{f}$ as the least-squares solution to a linear regression problem with responses $Y_1,\ldots,Y_n$ and features $v_1,\ldots,v_K$.
-Often, the Laplacian eigenmaps embedding is viewed as a tool for dimensionality reduction, wherein it is implicitly assumed that $K$ is much smaller than $d$.
-We will neither explicitly nor implicitly take $K < d$; indeed, the embedding perspective is not particularly illuminating in what follows, and we do not henceforth make reference to it.
-Instead, we use ``Laplacian eigenmaps'' to directly refer to the estimator $\wh{f}$ or test statistic $\wh{T}$.
-
 \subsection{Spectral Projection using Continuum Eigenfunctions}
 \label{subsec:spectral_projection}
-Laplacian Eigenmaps serves as a data-dependent alternative to more classical spectral projection methods.
-Classically, spectral projection methods operate by projecting the response vector ${\bf Y}$ onto a subspace spanned by eigenfunctions of some differential operator.
+Laplacian eigenmaps serves as a data-dependent alternative to more classical population-level spectral projection methods, which operate by projecting the response vector ${\bf Y}$ onto a subspace spanned by eigenfunctions of some differential operator.
 For instance, projecting ${\bf Y}$ onto the eigenfunctions of the Laplacian operator $\Delta = \sum_{i = 1}^{d} \partial^2f/\partial x_i^2$ corresponds to estimating $f_0$ using noisy empirical Fourier coefficients.
 
-We will focus on a density-dependent altnerative to the Laplacian,  $\Delta_P:C^2(\mc{X}) \to L^2(\mc{X})$,
+We will focus on a density-dependent alternative to the Laplacian,  $\Delta_P:C^2(\mc{X}) \to L^2(\mc{X})$,
 \begin{equation}
@@
 \end{equation}
-and sorted as usual in ascending order of eigenvalue.\footnote{For formal justification of why~\eqref{eqn:laplace_beltrami_eigenproblem} has discrete spectrum under either Model~\ref{def:model_flat_euclidean} or~\ref{def:model_manifold}, see~\cite{garciatrillos18,trillos2019}.} Formally, the classical spectral series estimator and test statistic are\footnote{Technically speaking, $\wt{f}$ is not the projection of ${\bf Y}$ onto the subspace of $L^2(P_n)$ spanned by $\psi_1,\ldots,\psi_K$, since $\dotp{\psi_k}{\psi_{\ell}}_n \neq 0$.
+and sorted as usual in ascending order of eigenvalue.\footnote{For formal justification of the fact that~\eqref{eqn:laplace_beltrami_eigenproblem} has a discrete spectrum under either Model~\ref{def:model_flat_euclidean} or~\ref{def:model_manifold}, see~\cite{garciatrillos18,trillos2019}.} Formally, the classical spectral series estimator and test statistic are\footnote{Technically speaking, $\wt{f}$ is not the projection of ${\bf Y}$ onto the subspace of $L^2(P_n)$ spanned by $\psi_1,\ldots,\psi_K$, since $\dotp{\psi_k}{\psi_{\ell}}_n \neq 0$.
 However, these eigenfunctions are very close to orthogonal in $L^2(P_n)$, orthogonalization will have a negligible impact on the overall statistical properties, and so for simplicity we consider the methods as defined in~\eqref{eqn:classical_spectral_projection}.}
@@
 \label{eqn:classical_spectral_projection}
-\wt{f}(x) := \frac{1}{n}\sum_{k = 1}^{K} \dotp{{\bf Y}}{\psi_k}_{n} \psi_k(x),\quad\textrm{and}\quad \wt{T} := \|\wt{f}\|_n^2.
+\wt{f}(x) := \sum_{k = 1}^{K} \dotp{{\bf Y}}{\psi_k}_{n} \psi_k(x),\quad\textrm{and}\quad \wt{T} := \|\wt{f}\|_P^2.
 \end{equation}
 
-We focus on the particular differential operator $\Delta_P$ because the spectrum of $\Delta_P$ is the population-level limit of graph Laplacian spectra: for any fixed $k \in \mathbb{N}$, as the number of samples $n \to \infty$ and the graph radius $\varepsilon \to 0$,
+We focus on the particular differential operator $\Delta_P$ because the spectrum of $\Delta_P$ is the population-level limit of the graph Laplacian spectrum: for any fixed $k \in \mathbb{N}$, as the number of samples $n \to \infty$, the graph radius $\varepsilon \to 0$, and $n\varepsilon^{2 + \dim(\mc{X})} \to \infty$,
 \begin{equation}
@@
 This holds in either the flat Euclidean (Model~\ref{def:model_flat_euclidean}) or manifold (Model~\ref{def:model_manifold}) setups~\citep{garciatrillos18,trillos2019}.
-It follows that for any fixed number of eigenvectors $K \in \mathbb{N}$, both $\wh{f} \to \wt{f}$ and $\wh{T} \to \wt{T}$, and in this way Laplacian Eigenmaps can be formally tied to classical spectral projection methods for regression.
+It follows that for any fixed number of eigenvectors $K \in \mathbb{N}$, both $\wh{f} \to \wt{f}$ and $\wh{T} \to \wt{T}$, and in this way Laplacian eigenmaps can be formally tied to more classical spectral projection methods for regression.
 
-On the other hand, the eigenvectors and eigenfunctions of $\Delta_P$ can also be used to give a spectral definition of Sobolev spaces on general domains $\mc{X}$. Consider the ellipsoid
+On the other hand, the eigenvalues and eigenfunctions of $\Delta_P$ can also be used to give a spectral definition of Sobolev spaces on general domains $\mc{X}$. Consider the ellipsoid
 \begin{equation}
 \label{eqn:sobolev_ellipsoid}
-\mc{H}^{s}(\mc{X}) := \Bigl\{\sum_{k = 1}^{\infty} a_k \psi_k \in L^2(\mc{X}):  \sum_{k = 1}^{\infty} a_k^2 \lambda_{k}(\Delta_P) \leq M^2 \Bigr\},
+\mc{H}^{s}(\mc{X}) := \Bigl\{\sum_{k = 1}^{\infty} a_k \psi_k \in L^2(\mc{X}):  \sum_{k = 1}^{\infty} a_k^2 \lambda_{k}^s(\Delta_P) \leq M^2 \Bigr\},
 \end{equation}
-equipped with the norm $\|\sum_{k = 1}^{\infty} a_k \psi_k\|_{\mc{H}^s(\mc{X})}^2 = \sum_{k = 1}^{\infty} a_k^2 \lambda_{k}(\Delta_P)$.
+equipped with the norm $\|\sum_{k = 1}^{\infty} a_k \psi_k\|_{\mc{H}^s(\mc{X})}^2 = \sum_{k = 1}^{\infty} a_k^2 \lambda_{k}^s(\Delta_P)$.
 Under appropriate regularity conditions $\mc{H}^s(\mc{X})$ consists of functions $f \in H^s(\mc{X})$ which also satisfy some additional boundary conditions.
@@
 
-This spectral formulation suggests that classical spectral projection methods---and in turn Laplacian Eigenmaps---are natural choices for regression over Sobolev spaces.
-Furthermore, since classical spectral projection methods serve, in a sense, as the infinite-data limit of Laplacian eigenmaps, the statistical properties of the former---which are generally speaking better understood and easier to derive---should intuitively shed some light on the properties of the latter.
-To this end, we now give a pair of results, Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, which establish that classical spectral projection methods are statistically optimal for nonparametric estimation and testing over Sobolev ellipsoids.
+This spectral formulation suggests that population-level spectral projection methods---and in turn Laplacian eigenmaps---are natural choices for regression over Sobolev spaces.
+Furthermore, since these population-level methods serve, in a sense, as the infinite-data limit of Laplacian eigenmaps, the statistical properties of the former---which are generally speaking better understood and easier to derive---should intuitively shed some light on the properties of the latter.
+To this end, we now state a pair of results, Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, which establish that population-level spectral projection methods are statistically optimal for nonparametric estimation and testing over Sobolev ellipsoids.
 These results are generalizations of previously known upper bounds on the estimation and testing error of classical spectral methods~\citep{tsybakov08,ingster2009}, but they hold under less stringent (nonparametric) conditions on the design distribution $P$.
-This latter point is important, since as we have argued one attractive quality of Laplacian Eigenmaps is its adaptivity to general design distributions.
+This latter point is important, since as we have argued one attractive quality of Laplacian eigenmaps is its adaptivity to general design distributions.
 
@@
 Suppose additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$, $f_0 \in \mc{H}^{s}(\mc{X};M)$ and $\|f_0\|_P^2 \leq 1$.
-Then there exists a constant $C$ which does not depend on $f_0,M$ or $n$ such that the following statement holds: if the spectral projection estimator $\wt{f}$ is computed with parameter $K = \floor{M^2n}^{d/(2s + d)}$, then
+Then there exists a constant $C$ which does not depend on $f_0$, $M$ or $n$ such that the following statement holds: if the spectral projection estimator $\wt{f}$ is computed with parameter $K = \max\{\floor{M^2n}^{d/(2s + d)},1\}$, then
 	\begin{equation}
 	\label{eqn:spectral_series_estimation}
-	\Ebb\bigl[\|\wt{f} - f_0\|_P^2\bigr] \leq C \min\Bigl\{M^2\bigl(M^2n\bigr)^{-2s/(2s + d)}, M^2\Bigr\}.
+	\Ebb\bigl[\|\wt{f} - f_0\|_P^2\bigr] \leq C \min\Bigl\{M^2\bigl(M^2n\bigr)^{-2s/(2s + d)}, \frac{1}{n}\Bigr\}.
 	\end{equation}
 \end{proposition}
-When the Sobolev ball radius $M \asymp 1$---which automatically implies $\|f_0\|_{P}^2 \lesssim 1$ as well---the upper bound in~\eqref{eqn:spectral_series_estimation} is on the order of $n^{-2s/(2s + d)}$, which matches the well-known minimax rate for estimation over Sobolev classes (see e.g.~\cite{wasserman2006,tsybakov08} and references therein).
+When the Sobolev ball radius $n^{-1/2} \lesssim M$, the upper bound in~\eqref{eqn:spectral_series_estimation} is on the order of $M^2(M^2n)^{-2s/(2s + d)}$, which is the well-known minimax rate for estimation over Sobolev classes (see e.g.~\cite{gyorfi2006,wasserman2006,tsybakov08} and references therein, and specifically Theorem~3.2 of~\cite{gyorfi2006} for a matching lower bound in the context of nonparametric regression with random design).
 
@@
 We now give the proof of Proposition~\ref{prop:spectral_series_estimation}.
-The structure of the analysis, which is fairly classical and straightforward, can be usefully compared to our analysis of Laplacian Eigenmaps (see Section~\ref{subsec:analysis}).
+The structure of the analysis, which is fairly classical and straightforward, can be usefully compared to our analysis of Laplacian eigenmaps (see Section~\ref{subsec:analysis}).
 
@@
 \end{equation}
-Equation~\eqref{eqn:local_weyl} follows straightforwardly from Theorem 17.5.3 of~\cite{hormander1973}.
-Plugging the upper bounds~\eqref{eqn:weyl} and~\eqref{eqn:local_weyl} back into~\eqref{pf:spectral_series_estimation_1} shows that
+Equation~\eqref{eqn:local_weyl} is a direct implication of~\eqref{eqn:weyl} along with Theorem 17.5.3 of~\cite{hormander1973}.
+Plugging the upper bounds~\eqref{eqn:weyl} and~\eqref{eqn:local_weyl} back into~\eqref{pf:spectral_series_estimation_1}, we conclude that
 \begin{equation}
 \label{pf:spectral_series_estimation_3}
-\Ebb \|\wt{f} - f_0\|_P^2 \leq C\biggl(\frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{(K + 1)^{2s/d}} + \frac{K}{n}\biggr),
+\Ebb \|\wt{f} - f_0\|_P^2 \leq C\biggl(\frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{(K + 1)^{2s/d}} + \frac{K}{n}\biggr).
 \end{equation}
-and setting $K = \floor{(1 + \|f_0\|_{P}^2)M^2n}^{d/(2s + d)}$, which balances squared bias and variance, then yields the claim.
+If $n^{-1/2} \geq M$, then taking $K = 1$ implies $\Ebb \|\wt{f} - f_0\|_P^2 \leq C(M^2 + 1/n)$.
+Otherwise, setting $K = \floor{M^2n}^{d/(2s + d)}$ balances squared bias and variance, and yields the claim.
 \qed.
@@
 \end{equation} 
-To fix ideas, here and throughout we focus on the signal detection problem, meaning the special case where $f_0^{\star} = 0$.\footnote{This is without loss of generality since all the test statistics we consider are easily modified to handle the case when $f_0^{\ast}$ is not $0$, by simply subtracting $f_0^{\ast}(X_i)$ from each observation $Y_i$, with no change in the analysis.}
+To fix ideas, here and throughout we focus on the signal detection problem, meaning the special case where $f_0^{\star} = 0$.\footnote{This is without loss of generality since all the test statistics we consider are easily modified to handle the case when $f_0^{\ast}$ is not $0$, by simply subtracting $f_0^{\ast}(X_i)$ from each observation $Y_i$, with no change in the analysis.} For more background on nonparametric goodness-of-fit testing problems, see~\cite{ingster2012}.
 
-In this case, the classical spectral projection test $\wt{\varphi} = \1\{\wt{T} \geq K/N + \sqrt{2K/an}\}$ has bounded Type I error, $\Ebb_{0}[\wt{\varphi}] \leq a (1 + o(1))$.
-Proposition~\ref{prop:spectral_series_testing} gives an upper bound on the Type II error that holds uniformly over all $f_0 \in \mc{H}^s(\mc{X};M)$ for which $\|f_0\|_P^2$ is sufficiently large.
+For the signal detection problem, the population-level spectral projection test $\wt{\varphi} := \1\{\wt{T} \geq K/N + \sqrt{2K/an^2}\}$ has bounded Type I error, $\Ebb_{0}[\wt{\varphi}] \leq a (1 + o(1))$.
+Proposition~\ref{prop:spectral_series_testing} gives an upper bound on the Type II error that holds over all $f_0 \in \mc{H}^s(\mc{X};M)$ for which $\|f_0\|_P^2$ is sufficiently large.
 \begin{proposition}
@@
 	Suppose data is observed according to Model~\ref{def:model_flat_euclidean}, and that the density $p$ is known.
-Suppose additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$, $f_0 \in \mc{H}^{s}(\mc{X};M)$ and $\|f_0\|_{\Leb^4(\mc{X})}^4 \leq 1$.
-Then there exists a constant $C$ which does not depend on $f_0,M$ or $n$ such that the following statement holds: if the spectral projection test $\wt{\varphi}$ is computed with parameter $K = \floor{M^2n}^{2d/(4s + d)}$, and if
+Suppose additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$, $f_0 \in \mc{H}^{s}(\mc{X};M)$ for some $s > d/4$, and $\|f_0\|_{\Leb^4(\mc{X})}^4 \leq 1$.
+Then there exists a constant $C$ which does not depend on $f_0$, $M$ or $n$ such that the following statement holds: if the spectral projection test $\wt{\varphi}$ is computed with parameter $K = \max\{\floor{M^2n}^{2d/(4s + d)},1\}$, and if
 	\begin{equation}
 	\label{eqn:spectral_series_testing}
-	\|f_0\|_P^2 \geq C\min\{M^2(M^2n)^{-4s/(4s + d)}, M^2\}
+	\|f_0\|_P^2 \geq C\min\Bigl\{M^2(M^2n)^{-4s/(4s + d)}, \frac{1}{n}\Bigr\}
 	\end{equation}
-	then the Type I error is bounded, $\Ebb_{f_0}[1 - \phi] \leq b$.
+	then the Type II error is upper bounded, $\Ebb_{f_0}[1 - \phi] \leq b$.
 \end{proposition}
-Focusing again on the special case where the Sobolev ball radius $M \asymp 1$---which automatically implies $\|f_0\|_{\Leb^4(\mc{X})}^4 \lesssim 1$ when $s > d/4$, due to a Sobolev embedding theorem---the upper bound in~\eqref{eqn:spectral_series_testing} is $n^{-4s/(4s + d)}$, matching the usual minimax critical radius over Sobolev spaces (for more details on goodness-of-fit testing see~\cite{ingster2009,ingster2012}).
-On the other hand, when $s \leq d/4$ the problem fundamentally changes in two ways.
-First, Sobolev space $H^s(\mc{X})$ no longer continuously embeds into $L^4(\mc{X})$, and indeed the minimax testing rates in this case are unknown.
-Second, if one explicitly assumes $f_0 \in L^4(\mc{X};1)$ then the minimax critical radius for regression testing is on the order of $n^{-1/2}$ \citep{guerre02}, and is achieved by a test using the naive statistic $\|{\bf Y}\|_n^2$.
+Assuming again that $n^{-1/2} \lesssim M$, the upper bound in~\eqref{eqn:spectral_series_testing} is $M^2(M^2n)^{-4s/(4s + d)}$, matching the usual minimax critical radius over Sobolev spaces (see e.g.~\cite{guerre02,ingster2009,ingster2012}.
+Specifically, \cite{ingster2009} show that the minimax squared critical radius is on the order of $n^{-4s/(4s + d)}$ when $M = 1$, and simple alterations of their analysis imply the rate $M^2(M^2n)^{-4s/(4s + d)}$ for general $M$.)
+
+On the other hand, when $s \leq d/4$ the minimax regression testing rates over $H^s(\mc{X})$ are not known.
+If one explicitly assumes $f_0 \in L^4(\mc{X};1)$---note that $H^s(\mc{X})$ does not continuously embed into $L^4(\mc{X})$ when $s \leq d/4$---then the minimax critical radius for regression testing is on the order of $n^{-1/2}$ \citep{guerre02}, and is achieved by a test using the naive statistic $\|{\bf Y}\|_n^2$.
 In other words, the regression testing problem over Sobolev spaces fundamentally changes when $s \leq d/4$, and hereafter when we discuss testing we will limit our consideration to $s > d/4$.
 \paragraph{Proof of Proposition~\ref{prop:spectral_series_testing}.}
-We briefly lay out the main ideas needed to prove Proposition~\ref{prop:spectral_series_testing}, following the lead of~\cite{ingster2009} who prove a similar result in the special case where $P$ is the uniform distribution over $\mc{X} = [0,1]^d$, and referring to that work for more details.
+We briefly lay out the main ideas needed to prove Proposition~\ref{prop:spectral_series_testing}, following the lead of~\cite{ingster2009} who prove a similar result in the special case where $M = 1$ and $P$ is the uniform distribution over $\mc{X} = [0,1]^d$, and referring to that work for more details.
 
@@
 \begin{align*}
-\Ebb[\wt{T}] = \frac{(n - 1)}{n} \sum_{k = 1}^{K} \dotp{f_0}{\psi_K}_P^2 + \frac{K}{n} + \Ebb[(f_0(X))^2 \dot \sum_{k = 1}^{K} (\psi_k(X))^2],
+\Ebb[\wt{T}] = \frac{(n - 1)}{n} \sum_{k = 1}^{K} \dotp{f_0}{\psi_K}_P^2 + \frac{K}{n} + \Ebb\Bigl[(f_0(X))^2 \dot \sum_{k = 1}^{K} (\psi_k(X))^2\Bigr],
 \end{align*}
-from which we clearly see that under the alternative,
+and from~\eqref{eqn:weyl} (Weyl's Law) we have that under the alternative $f_0 \neq 0$,
 \begin{equation*}
-\Ebb_{f_0}[\wt{T}] \geq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\lambda_{K + 1}(\Delta_{P})^s} + K/n.
+\Ebb_{f_0}[\wt{T}] \geq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\lambda_{K + 1}^s(\Delta_{P})} + \frac{K}{n}.
 \end{equation*}
-To compute the variance, we decompose $\wt{T} = \wt{T}_{1,1} + \wt{T}_{1,2} + \wt{T}_{1,3} + \wt{T}_2$ into the sum of 3 U-statistics and the remaining diagonal terms; letting $\kappa(x,x') = \sum_{k = 1}^{K} \psi_k(x) \psi_k(x')$ be the equivalent kernel, we have that
+To compute the variance, we decompose $\wt{T} = \wt{T}_{1,1} + \wt{T}_{1,2} + \wt{T}_{1,3} + \wt{T}_2$ into the sum of 3 U-statistics and the remaining diagonal terms, defined in terms of the equivalent kernel $\kappa(x,x') = \sum_{k = 1}^{K} \psi_k(x) \psi_k(x')$ as,
 \begin{align*}
-T_{1,1} & = \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} w_i w_j \kappa(X_i,X_j),\quad && T_{1,2} = \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} \bigl(w_i f_0(X_j) + w_jf_0(X_i)\bigr)\kappa(X_i,X_j) \\
-T_{1,3} & = \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} f_0(X_i) f_0(X_j) \kappa(X_i,X_j), \quad && T_2 = \frac{1}{n^2} \sum_{i = 1}^{n} Y_i^2 \kappa(X_i,X_i).
+T_{1,1} & := \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} w_i w_j \kappa(X_i,X_j),\quad && T_{1,2} := \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} \bigl(w_i f_0(X_j) + w_jf_0(X_i)\bigr)\kappa(X_i,X_j) \\
+T_{1,3} & := \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} f_0(X_i) f_0(X_j) \kappa(X_i,X_j), \quad && T_2 := \frac{1}{n^2} \sum_{i = 1}^{n} Y_i^2 \kappa(X_i,X_i).
 \end{align*}
@@
 \end{align*}
-where $\mathrm{(i)}-\mathrm{(iii)}$ hold due to local Weyl's law, meaning the upper bound on $\sum_{k = 1}^{K} (\psi_k(x))^2$ given in~\eqref{eqn:local_weyl}. Upper bounds on Type I and Type II error,
+where $\mathrm{(i)}-\mathrm{(iii)}$ hold due to local Weyl's law, i.e.~\eqref{eqn:local_weyl}. Upper bounds on Type I and Type II error,
 \begin{equation*}
@@
 \label{pf:spectral_series_test}
-\|f_0\|_P^2 \geq C\Biggl(\frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\lambda_{K + 1}(\Delta_{P})^s} + \sqrt{\frac{2K}{n}}\biggl(\sqrt{\frac{1}{a}} + \sqrt{\frac{1}{b}}\biggr)\Biggr)
+\|f_0\|_P^2 \geq C\Biggl(\frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\lambda_{K + 1}^s(\Delta_{P})} + \frac{\sqrt{K}}{n}\biggl(\sqrt{\frac{1}{a}} + \sqrt{\frac{1}{b}}\biggr)\Biggr)
 \end{equation}
 for a sufficiently large constant $C$, then $\Ebb_{f_0}[1 - \wt{\varphi}] \leq b$.
-The two summands in~\eqref{pf:spectral_series_test} are bias and standard deviation terms, respectively; choosing $K = \floor{M^2n}^{2d/(4s + d)}$ balances these two terms, and leads to~\eqref{eqn:spectral_series_testing}.
+The two summands in~\eqref{pf:spectral_series_test} are bias and standard deviation terms, respectively.
+When $M^2 \leq n^{-1}$, setting $K = 1$ gives the desired result.
+Otherwise, choosing $K = \floor{M^2n}^{2d/(4s + d)}$ balances these two terms, and leads to~\eqref{eqn:spectral_series_testing}.
 \qed
@@
 Of course, it is worth pointing out that other methods besides Laplacian eigenmaps are statistically optimal for nonparametric regression even when $p$ is unknown.
-We comment more on some of these in Section~\ref{sec:discussion}, after we have derived our major results regarding Laplacian Eigenmaps.
+We comment more on some of these in Section~\ref{sec:discussion}, after we have derived our major results regarding Laplacian eigenmaps.

### 2021-09-04 manuscripts/laplacian_eigenmaps/annals_of_statistics/laplacian_eigenmaps_regression_aos.tex
Put Laplacian Eigenmaps paper into AOS format.

--- 2021-08-31 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
@@
-\documentclass{article}
-\usepackage{amsmath}
-\usepackage{amsfonts, amsthm, amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\hypersetup{
-	colorlinks=true,
-	linkcolor=blue,
-	citecolor=blue
+\documentclass[aos]{imsart}
+
+%% Packages
+\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
+% \RequirePackage[numbers]{natbib}
+\RequirePackage[authoryear]{natbib}
+\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
+\RequirePackage{graphicx}
+\usepackage{mathtools, enumitem}
+\usepackage{multirow}
+
+\startlocaldefs
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% For Axiom, Claim, Corollary, Hypothesis, %%
+%% Lemma, Theorem, Proposition              %%
+%% use \theoremstyle{plain}                 %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
+\newtheorem{corollary}{Corollary}
+\newtheorem{proposition}{Proposition}
+\theoremstyle{definition}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% For Assumption, Definition, Example,     %%
+%% Notation, Property, Remark, Fact         %%
+%% use \theoremstyle{remark}                %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\theoremstyle{remark}
+\newtheorem{remark}{Remark}
+\newtheorem{definition}{Definition}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Please put your definitions here:        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%% Begin Ryan's definitions
+
+% Widebar
+\makeatletter
+\newcommand*\rel@kern[1]{\kern#1\dimexpr\macc@kerna}
+\newcommand*\widebar[1]{%
+	\begingroup
+	\def\mathaccent##1##2{%
+		\rel@kern{0.8}%
+		\overline{\rel@kern{-0.8}\macc@nucleus\rel@kern{0.2}}%
+		\rel@kern{-0.2}%
+	}%
+	\macc@depth\@ne
+	\let\math@bgroup\@empty \let\math@egroup\macc@set@skewchar
+	\mathsurround\z@ \frozen@everymath{\mathgroup\macc@group\relax}%
+	\macc@set@skewchar\relax
+	\let\mathaccentV\macc@nested@a
+	\macc@nested@a\relax111{#1}%
+	\endgroup
 }
-\usepackage[parfill]{parskip}
-\usepackage{algpseudocode}
-\usepackage{algorithm}
-\usepackage{enumerate}
-\usepackage[shortlabels]{enumitem}
-\usepackage{fullpage}
-\usepackage{mathtools}
-\usepackage{tikz}
-\usepackage{bm}
-\usepackage{multirow}
-\usepackage[font={small,it}]{caption}
-
-\usepackage{natbib}
-\renewcommand{\bibname}{References}
-\renewcommand{\bibsection}{\subsubsection*{\bibname}}
+\makeatother
+
+% Min and max
+\newcommand{\argmin}{\mathop{\mathrm{argmin}}}
+\newcommand{\argmax}{\mathop{\mathrm{argmax}}}
+\newcommand{\minimize}{\mathop{\mathrm{minimize}}}
+\newcommand{\st}{\mathop{\mathrm{subject\,\,to}}}
+
+% Shortcuts
+\def\R{\mathbb{R}}
+
+%%% End Ryan's definitions
+
+%%% Begin Alden's additions
+\newcommand{\Ebb}{\mathbb{E}}
+\newcommand{\Pbb}{\mathbb{P}}
+\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
+\newcommand{\wt}[1]{\widetilde{#1}}
+\newcommand{\wh}[1]{\widehat{#1}}
+\newcommand{\mc}[1]{\mathcal{#1}}
+\newcommand{\Reals}{\mathbb{R}} % Same thing as Ryan's \R
+\newcommand{\Rd}{\Reals^d}
+\newcommand{\wb}[1]{\widebar{#1}}
+\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
+\newcommand{\Var}{\mathrm{Var}}
+\newcommand{\Cov}{\mathrm{Cov}}
+\newcommand{\1}{\mathbf{1}}
+\newcommand{\bj}{{\bf j}}
+\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
 
@@
 \DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
-\DeclareMathAccent{\wb}{0}{mathx}{"73}
 \DeclareMathAccent{\wc}{0}{mathx}{"71}
-
-\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
-\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}
-
-\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
-\newcommand{\Graph}{\mathcal{G}}
-\newcommand{\Reals}{\mathbb{R}}
-\newcommand{\Identity}{\mathbb{I}}
-\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
-\newcommand{\convprob}{\overset{p}{\to}}
-\newcommand{\convdist}{\overset{w}{\to}}
-\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
-\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
-\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
-\newcommand{\iset}{\mathbf{i}}
-\newcommand{\jset}{\mathbf{j}}
-\newcommand{\myexp}[1]{\exp \{ #1 \}}
-\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
-\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
-\newcommand{\ext}[1]{\widetilde{#1}}
-\newcommand{\set}[1]{\left\{#1\right\}}
-\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
-\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
-\newcommand{\Var}{\mathrm{Var}}
-\newcommand{\Cov}{\mathrm{Cov}}
-\newcommand{\diam}{\mathrm{diam}}
-
-\newcommand{\emC}{C_n}
-\newcommand{\emCpr}{C'_n}
-\newcommand{\emCthick}{C^{\sigma}_n}
-\newcommand{\emCprthick}{C'^{\sigma}_n}
-\newcommand{\emS}{S^{\sigma}_n}
-\newcommand{\estC}{\widehat{C}_n}
-\newcommand{\hC}{\hat{C^{\sigma}_n}}
-\newcommand{\vol}{\text{vol}}
-\newcommand{\spansp}{\mathrm{span}~}
-\newcommand{\1}{\mathbf{1}}
-
-\newcommand{\Linv}{L^{\Xsetagger}}
-\DeclareMathOperator*{\argmin}{argmin}
-\DeclareMathOperator*{\argmax}{argmax}
-
-\newcommand{\emF}{\mathbb{F}_n}
-\newcommand{\emG}{\mathbb{G}_n}
-\newcommand{\emP}{\mathbb{P}_n}
-\newcommand{\F}{\mathcal{F}}
-\newcommand{\D}{\mathcal{D}}
-\newcommand{\R}{\mathcal{R}}
-\newcommand{\Rd}{\Reals^d}
-\newcommand{\RD}{\Reals^D}
-\newcommand{\Nbb}{\mathbb{N}}
-
-%%% Vectors
-\newcommand{\thetast}{\theta^{\star}}
-\newcommand{\betap}{\beta^{(p)}}
-\newcommand{\betaq}{\beta^{(q)}}
-\newcommand{\vardeltapq}{\varDelta^{(p,q)}}
-\newcommand{\lambdavec}{\boldsymbol{\lambda}}
-\newcommand{\bj}{{\bf j}}
-
-%%% Matrices
-\newcommand{\X}{X} % no bold
-\newcommand{\Y}{Y} % no bold
-\newcommand{\Z}{Z} % no bold
-\newcommand{\Lgrid}{L_{\grid}}
-\newcommand{\Xsetgrid}{D_{\grid}}
-\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}
-\newcommand{\Lap}{{\bf L}}
-\newcommand{\NLap}{{\bf N}}
-\newcommand{\PLap}{{\bf P}}
-\newcommand{\Id}{I}
-
-%%% Sets and classes
-\newcommand{\Xset}{\mathcal{X}}
-\newcommand{\Sset}{\mathcal{S}}
-\newcommand{\Hclass}{\mathcal{H}}
-\newcommand{\Pclass}{\mathcal{P}}
-\newcommand{\Leb}{L}
-\newcommand{\mc}[1]{\mathcal{#1}}
-
-%%% Distributions and related quantities
-\newcommand{\Pbb}{\mathbb{P}}
-\newcommand{\Ebb}{\mathbb{E}}
-\newcommand{\Qbb}{\mathbb{Q}}
-\newcommand{\Ibb}{\mathbb{I}}
-
-%%% Operators
-\newcommand{\Tadj}{T^{\star}}
-\newcommand{\Xsetive}{\mathrm{div}}
-\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
-\newcommand{\gradient}{\mathcal{D}}
-\newcommand{\Hessian}{\mathcal{D}^2}
-\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
-\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}
-
-%%% Misc
-\newcommand{\grid}{\mathrm{grid}}
-\newcommand{\critr}{R_n}
-\newcommand{\Xsetx}{\,dx}
-\newcommand{\Xsety}{\,dy}
-\newcommand{\Xsetr}{\,dr}
-\newcommand{\Xsetxpr}{\,dx'}
-\newcommand{\Xsetypr}{\,dy'}
-\newcommand{\wt}[1]{\widetilde{#1}}
-\newcommand{\wh}[1]{\widehat{#1}}
-\newcommand{\ol}[1]{\overline{#1}}
-\newcommand{\spec}{\mathrm{spec}}
-\newcommand{\LE}{\mathrm{LE}}
-\newcommand{\LS}{\mathrm{LS}}
-\newcommand{\SM}{\mathrm{SM}}
-\newcommand{\OS}{\mathrm{FS}}
-\newcommand{\PLS}{\mathrm{PLS}}
-
-%%% Order of magnitude
-\newcommand{\soom}{\sim}
-
-% \newcommand{\span}{\textrm{span}}
-
-\newtheoremstyle{alden}
-{6pt} % Space above
-{6pt} % Space below
-{} % Body font
-{} % Indent amount
-{\bfseries} % Theorem head font
-{.} % Punctuation after theorem head
-{.5em} % Space after theorem head
-{} % Theorem head spec (can be left empty, meaning `normal')
-
-\theoremstyle{alden}
-
-\newtheoremstyle{aldenthm}
-{6pt} % Space above
-{6pt} % Space below
-{\itshape} % Body font
-{} % Indent amount
-{\bfseries} % Theorem head font
-{.} % Punctuation after theorem head
-{.5em} % Space after theorem head
-{} % Theorem head spec (can be left empty, meaning `normal')
-
-\theoremstyle{aldenthm}
-\newtheorem{theorem}{Theorem}
-\newtheorem{conjecture}{Conjecture}
-\newtheorem{lemma}{Lemma}
-\newtheorem{example}{Example}
-\newtheorem{corollary}{Corollary}
-\newtheorem{proposition}{Proposition}
-\newtheorem{assumption}{Assumption}
-\newtheorem{remark}{Remark}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-
-\theoremstyle{remark}
+%%% End Alden's Additions
+
+\endlocaldefs
 
 \begin{document}
-\title{Minimax-optimal Laplacian Eigenmaps regression over Sobolev Spaces with Neighborhood Graphs}
-\author{Alden Green}
-\date{\today}
-\maketitle
-
+
+\begin{frontmatter}
+
+\title{Minimax-optimal Regression over Sobolev Spaces via Laplacian Eigenmaps with Neighborhood Graphs}
+\runtitle{Minimax-optimal Laplacian Eigenmaps Regression}
+%\thankstext{T1}{A sample of additional note to the title.}
+
+\begin{aug}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Only one address is permitted per author. %%
+%% Only division, organization and e-mail is %%
+%% included in the address.                  %%
+%% Additional information can be included in %%
+%% the Acknowledgments section if necessary. %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\author[A]{\fnms{Alden} \snm{Green}\ead[label=e1]{}},
+\author[A]{\fnms{Sivaraman} \snm{Balakrishnan}\ead[label=e1]{}}
+\and
+\author[A]{\fnms{Ryan J.} \snm{Tibshirani}\ead[label=e1]{}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Addresses                                %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\address[A]{Carnegie Mellon University, \printead{e1}}
+\end{aug}
+
+\begin{abstract}
+In this paper we study the statistical properties of nonparametric regression methods based on \emph{Laplacian eigenmaps}.
+These are spectral projection methods, which involve projecting a vector of observed responses onto a subspace spanned by certain eigenvectors of a neighborhood graph Laplacian.
+We show that in many cases these methods achieve minimax rates of convergence over continuum function classes.
+Specifically, we consider the problem of random design regression over Sobolev spaces $H^s(\mc{X})$---where $\mc{X} \subseteq \Rd$ is the support of the design density $p$---and show that under sufficient smoothness conditions on $p$, methods based on Laplacian eigenmaps achieve the optimal rates for both estimation (where the optimal rate is known to be $n^{-2s/(2s + d)}$) and goodness-of-fit testing ($n^{-4s/(4s + d)}$).
+We also show that Laplacian eigenmaps is \emph{manifold adaptive}: that is, we consider the situation where $\mc{X}$ is a manifold of small intrinsic dimension $m$, and give upper bounds establishing that Laplacian eigenmaps achieves the faster minimax estimation ($n^{-2s/(2s + m)}$) and testing ($n^{-4s/(4s + m)}$) rates of convergence.
+Finally, since the Laplacian eigenmaps estimator is defined only at observed design points, we propose an out-of-sample extension based on kernel smoothing, and prove that it achieves optimal out-of-sample error.
+We support these theoretical results with empirical evidence.
+\end{abstract}
+
+\begin{keyword}[class=MSC]
+\kwd{62G08, 62G10, 68R10}
+\end{keyword}
+
+\begin{keyword}
+\kwd{Laplacian eigenmaps, nonparametric regression, Sobolev space, minimax convergence rate}
+\end{keyword}
+
+\end{frontmatter}
+% Main text
 \section{Introduction}
@@
 More specifically, we consider the worst-case risk of Laplacian eigenmaps over normed balls in Sobolev spaces.
-Roughly speaking, for a domain $\mc{X} \subseteq \Rd$, the order-$s$ (Hilbert-)Sobolev space $H^s(\mc{X})$ consists of those functions which are $s$-times weakly differentiable, with order-$1$ up through order-$s$ derivatives bounded in $L^2(\mc{X})$ norm.
+Roughly speaking, for a domain $\mc{X} \subseteq \Rd$, the order-$s$ (Hilbert)-Sobolev space $H^s(\mc{X})$ consists of those functions which are $s$-times weakly differentiable, with order-$1$ up through order-$s$ derivatives bounded in $L^2(\mc{X})$ norm.
 Equivalently, Sobolev spaces can be defined in a spectral manner;  letting $\psi_1,\psi_2,\ldots$ be the eigenfunctions of a Laplacian operator defined with respect to $\mc{X}$, Sobolev spaces consist of those functions $f$ whose Fourier coefficients $\theta_k := \int_{\mc{X}} f \psi_k$ are bounded in a particular sequence norm (see Section~\ref{subsec:spectral_projection}).
@@
 
-To derive these results, we develop a framework for analysis reminiscent of that  classically used to analyze population-level spectral projection estimators, but ultimately quite different.
+To derive these results, we develop a framework for analysis reminiscent of that  classically used to analyze population-level spectral projection methods, but ultimately quite different.
 The chief difference is in the approximation error (bias).
@@
 			\hline
-			$s \leq 3$ & $\bm{n^{-2s/(2s + d)}}$ & $\bm{n^{-2s/(2s + m)}}$ \\
-			$s > 3$  & $\bm{n^{-2s/(2s + d)}}$ & $n^{-6/(6 + m)}$
+			$s \leq 3$ & ${\bf n^{-2s/(2s + d)}}$ & ${\bf n^{-2s/(2s + m)}}$ \\
+			$s > 3$  & ${\bf n^{-2s/(2s + d)}}$ & $n^{-6/(6 + m)}$
 		\end{tabular}
@@
 			\hline
-			\multirow{2}{*}{$s = 1$} & $\dim(\mc{X}) < 4$ & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
-			& $\dim(\mc{X}) \geq 4$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
+			\multirow{2}{*}{$s = 1$} & $\dim(\mc{X}) < 4$ & ${\bf n^{-4s/(4s + d)}}$ & ${\bf n^{-4s/(4s + m)}}$ \\
+			& $\dim(\mc{X}) \geq 4$ & ${\bf n^{-1/2}}$ & ${\bf n^{-1/2}}$ \\
 			\hline
-			\multirow{3}{*}{$s = 2$ or $3$} & $\dim(\mc{X}) \leq 4$  & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
+			\multirow{3}{*}{$s = 2$ or $3$} & $\dim(\mc{X}) \leq 4$  & ${\bf n^{-4s/(4s + d)}}$ & ${\bf n^{-4s/(4s + m)}}$ \\
 			& $4 <\dim(\mc{X}) < 4s$  & $n^{-2s/(2(s - 1) + d)}$ & $n^{-2s/(2(s - 1) + m)}$\\
-			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
+			& $\dim(\mc{X}) \geq 4s$ & ${\bf n^{-1/2}}$ & ${\bf n^{-1/2}}$ \\
 			\hline
-			\multirow{3}{*}{$s > 3$} & $\dim(\mc{X}) \leq 4$ & $\bm{n^{-4s/(4s + d)}}$ & $n^{-12/(12 + d)}$ \\
+			\multirow{3}{*}{$s > 3$} & $\dim(\mc{X}) \leq 4$ & ${\bf n^{-4s/(4s + d)}}$ & $n^{-12/(12 + d)}$ \\
 			& $4 < \dim(\mc{X}) < 4s$ & $n^{-2s/(2(s - 1) + d)}$ & $n^{-6/(4 + m)}$ \\
-			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
+			& $\dim(\mc{X}) \geq 4s$ & ${\bf n^{-1/2}}$ & ${\bf n^{-1/2}}$ \\
 		\end{tabular}
@@
 Rates when $d > 4s$ assume that $f_0 \in L^4(\mc{X})$, and depend on $\|f_0\|_{L^4(\mc{X})}$.
-Although we suppress it for simplicity, in all cases when othe Laplacian eigenmaps test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal.}
+Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal.}
 	\label{tbl:testing_rates}
@@
 
-\paragraph{Related work.}
+\subsection{Related work}
 
@@
 
-\paragraph{Roadmap.}
+\subsection{Roadmap}
 We now outline the structure of the rest of this paper.
@@
 
-\paragraph{Notation.}
+\subsection{Notation}
 We frequently refer to various classical function classes.
@@
 We define $\dotp{f}{g}_P := \int fg\,dP$, and let $L^2(P)$ contain those functions $f$ for which $\|f\|_P^2 := \dotp{f}{f}_P$ is finite.
-Equivalently, we let $L^2(P_n)$ consist of those ``functions'' $f: \set{X_1,\ldots,X_n} \to \Reals$ for which the empirical norm $\|f\|_{n}^2 := \frac{1}{n}\sum_{i = 1}^{n} \bigl(f(X_i)\bigr)^2 < \infty$.
+Equivalently, we let $L^2(P_n)$ consist of those ``functions'' $f: \{X_1,\ldots,X_n\} \to \Reals$ for which the empirical norm $\|f\|_{n}^2 := \frac{1}{n}\sum_{i = 1}^{n} \bigl(f(X_i)\bigr)^2 < \infty$.
 When there is no chance of confusion, we will sometimes associate functions in $L^2(P_n)$ with vectors in $\Reals^n$, and vice versa.
@@
 
-\paragraph{Flat Euclidean model.}
+\subsubsection{Flat Euclidean model}
 In Definitions~\ref{def:model_flat_euclidean}-\ref{def:zero_trace_sobolev_space}, we collect the assumptions we make when working under the flat Euclidean model.
@@
 	For an integer $s \geq 1$, a function $f \in L^2(\mc{X})$ belongs to the Sobolev space $H^s(\mc{X})$ if for all $|\alpha| \leq s$, the weak derivatives $D^{\alpha}f$ exist and satisfy $D^{\alpha}f \in L^2(\mc{X})$.
-The $j$th order semi-norm for $f \in H^s(\mc{X})$ is $|f|_{H^j(\mc{X})} := \sum_{|\alpha| = j}\|D^{\alpha}f\|_{\Leb^2(\mc{X})}$, and the corresponding norm
-	\begin{equation*}
-	\|f\|_{H^s(\mc{X})}^2 := \|f\|_{\Leb^2(\mc{X})}^2 + \sum_{j = 1}^{s} |f|_{H^j(\mc{X})}^2,
+The $j$th order semi-norm for $f \in H^s(\mc{X})$ is $|f|_{H^j(\mc{X})} := \sum_{|\alpha| = j}\|D^{\alpha}f\|_{L^2(\mc{X})}$, and the corresponding norm
+	\begin{equation*}
+	\|f\|_{H^s(\mc{X})}^2 := \|f\|_{L^2(\mc{X})}^2 + \sum_{j = 1}^{s} |f|_{H^j(\mc{X})}^2,
 	\end{equation*}
@@
 (Here $\partial/(\partial {\bf n})$ is the partial derivative operator in the direction of the normal vector $\mathbf{n}$.)
-\paragraph{Manifold model.}
+
+\subsubsection{Manifold model}
 As in the flat Euclidean case, we start with some regularity conditions on the design. One such condition will be on the reach $R$ of the manifold $\mc{X}$, which we recall is defined as follows:
@@
 	\label{def:sobolev_space_manifold}
-	A function $f \in \Leb^2(\mc{X})$ belongs to the Sobolev space $H^{s}(\mc{X})$ if for all $\abs{\alpha} \leq s$, the weak derivatives $D^{\alpha}f$ exist and satisfy  $D^{\alpha}f \in \Leb^2(\mc{X})$.
+	A function $f \in L^2(\mc{X})$ belongs to the Sobolev space $H^{s}(\mc{X})$ if for all $|\alpha| \leq s$, the weak derivatives $D^{\alpha}f$ exist and satisfy  $D^{\alpha}f \in L^2(\mc{X})$.
 The $j$th order semi-norm $|f|_{H^j(\mc{X})}$, the norm $\|f\|_{H^s(\mc{X})}$, and the ball $H^s(\mc{X};M)$ are all defined as in Definition~\ref{def:sobolev_space}.
@@
 
-\paragraph{Estimation.}
+\subsubsection{Estimation}
 We begin with an upper bound on the $L^2(P)$ risk of $\wt{f}$.
@@
 
-\paragraph{Proof of Proposition~\ref{prop:spectral_series_estimation}.}
+\begin{proof}{Proof of Proposition~\ref{prop:spectral_series_estimation}.}
 We decompose risk into squared bias and variance,
@@
 Otherwise, setting $K = \floor{M^2n}^{d/(2s + d)}$ balances squared bias and variance, and yields the claim.
-\qed.
-
-\paragraph{Testing.}
+\end{proof}
+
+\subsubsection{Testing}
 In the goodness-of-fit testing problem, one asks for a test function---formally, a Borel measurable function $\phi$ that takes values in $\{0,1\}$--- which can distinguish between the hypotheses
@@
 	Suppose data is observed according to Model~\ref{def:model_flat_euclidean}, and that the density $p$ is known.
-Suppose additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$, $f_0 \in \mc{H}^{s}(\mc{X};M)$ for some $s > d/4$, and $\|f_0\|_{\Leb^4(\mc{X})}^4 \leq 1$.
+Suppose additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$, $f_0 \in \mc{H}^{s}(\mc{X};M)$ for some $s > d/4$, and $\|f_0\|_{L^4(\mc{X})}^4 \leq 1$.
 Then there exists a constant $C$ which does not depend on $f_0$, $M$ or $n$ such that the following statement holds: if the spectral projection test $\wt{\varphi}$ is computed with parameter $K = \max\{\floor{M^2n}^{2d/(4s + d)},1\}$, and if
@@
 In other words, the regression testing problem over Sobolev spaces fundamentally changes when $s \leq d/4$, and hereafter when we discuss testing we will limit our consideration to $s > d/4$.
-\paragraph{Proof of Proposition~\ref{prop:spectral_series_testing}.}
+\begin{proof}{Proof of Proposition~\ref{prop:spectral_series_testing}.}
 We briefly lay out the main ideas needed to prove Proposition~\ref{prop:spectral_series_testing}, following the lead of~\cite{ingster2009} who prove a similar result in the special case where $M = 1$ and $P$ is the uniform distribution over $\mc{X} = [0,1]^d$, and referring to that work for more details.
@@
 Otherwise, choosing $K = \floor{M^2n}^{2d/(4s + d)}$ balances these two terms, and leads to~\eqref{eqn:spectral_series_testing}.
-\qed
+\end{proof}
 % (AG 8/19/21): Same comment as above, regarding the role of M in random design regression testing.
@@
 
-\paragraph{Estimation.} Laplacian eigenmaps depends on the kernel $\eta$ and two tuning parameters, the graph radius $\varepsilon$ and number of eigenvectors $K$.
+\subsubsection{Estimation} Laplacian eigenmaps depends on the kernel $\eta$ and two tuning parameters, the graph radius $\varepsilon$ and number of eigenvectors $K$.
 We will need to make some assumptions on each.
@@
 We comment on these assumptions after stating our first main theorem, regarding the estimation error of Laplacian eigenmaps.
-The proof of this theorem, along with the proofs of all subsequent results, can be found in the Appendix.
+The proof of this theorem, along with the proofs of all subsequent results, can be found in the Supplementary Material.
 \begin{theorem}
@@
 
-\paragraph{Testing.} Consider the test $\varphi = \1\{\wh{T} \geq t_{a}\}$, where $t_{a}$ is the threshold
+\subsubsection{Testing} Consider the test $\varphi = \1\{\wh{T} \geq t_{a}\}$, where $t_{a}$ is the threshold
 \begin{equation*}
@@
 This implies that $\varphi$ is a minimax rate-optimal test over $H^1(\mc{X};M)$ when $d \in \{1,2,3\}$.
-As mentioned previously, when $d \geq 4$ the first order Sobolev space $H^1(\mc{X})$ does not continuously embed into $\Leb^4(\mc{X})$, and in this case the optimal rates for regression testing over Sobolev spaces are unknown.
+As mentioned previously, when $d \geq 4$ the first order Sobolev space $H^1(\mc{X})$ does not continuously embed into $L^4(\mc{X})$, and in this case the optimal rates for regression testing over Sobolev spaces are unknown.
 
@@
 
-\paragraph{Estimation.}
+\subsubsection{Estimation}
 In order to show that $\wh{f}$ is an optimal estimator over $H_0^s(\mc{X};M)$, we will require that $\varepsilon$ be meaningfully larger than the lower bound in~\ref{asmp:parameters_estimation_fo}.
@@
 As a result, for certain versions of the nonparametric regression problem---e.g.
-when loss is measured in $\Leb^{\infty}$ norm, or when the design points $\{X_1,\ldots,X_n\}$ are assumed to be fixed---in a minimax sense even consistent estimation is not possible.
+when loss is measured in $L^{\infty}$ norm, or when the design points $\{X_1,\ldots,X_n\}$ are assumed to be fixed---in a minimax sense even consistent estimation is not possible.
 Likewise, certain estimators are ``off the table'', most notably RKHS-based methods such as thin-plate splines of degree $k \leq d/2$.
-Nevertheless, for random design regression with error measured in squared $\Leb^2(P)$-norm, the spectral projection estimator $\wt{f}$ obtains the standard minimax rates $n^{-2s/(2s + d)}$ for all values of $s$ and $d$.
-Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and \ref{thm:laplacian_eigenmaps_estimation_ho} show that the same is true with respect to Laplacian eigenmaps, when error is measured in squared $\Leb^2(P_n)$-norm.
+Nevertheless, for random design regression with error measured in squared $L^2(P)$-norm, the spectral projection estimator $\wt{f}$ obtains the standard minimax rates $n^{-2s/(2s + d)}$ for all values of $s$ and $d$.
+Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and \ref{thm:laplacian_eigenmaps_estimation_ho} show that the same is true with respect to Laplacian eigenmaps, when error is measured in squared $L^2(P_n)$-norm.
 	\item \emph{Smoothness of design density}.
@@
 As compared to a sparse $G$, this results in more non-zero entries in the graph Laplacian, and increases the computational burden involved in computing $\wh{f}$.
-To address this issue, in Appendix~\ref{sec:computational_considerations} we review some approaches to \emph{spectral sparsification}, in which one efficiently computes a sparse graph $\wc{G}$ that approximates $G$ in a spectral sense.
+To address this issue, in Appendix~H of the Supplementary Material we review some approaches to \emph{spectral sparsification}, in which one efficiently computes a sparse graph $\wc{G}$ that approximates $G$ in a spectral sense.
 The hope is that the Laplacian eigenmaps estimator $\wc{f}$ computed with respect to the sparsified graph $\wc{G}$ has similar statistical properties as $\wh{f}$, while being much faster to compute.
@@
 
-\paragraph{Testing.} The test $\varphi$ can adapt to the higher-order smoothness of $f_0$, when $\varepsilon$ and $K$ are chosen correctly.
+\subsubsection{Testing} The test $\varphi$ can adapt to the higher-order smoothness of $f_0$, when $\varepsilon$ and $K$ are chosen correctly.
 \begin{enumerate}[label=(P\arabic*)]
@@
 
-\paragraph{Estimates of graph Sobolev seminorms.}
+\subsubsection{Estimates of graph Sobolev seminorms}
 In Proposition~\ref{prop:graph_seminorm_fo} we restate an upper bound on the Dirichlet energy $\dotp{L_{n,\varepsilon}f}{f}_n$ from~\cite{green2021}. 
@@
 
-\paragraph{Neighborhood graph eigenvalues.}
+\subsubsection{Neighborhood graph eigenvalues}
 On the other hand, several recent works \citep{burago2014,garciatrillos18,calder2019} have analyzed the convergence of $\lambda_{k}$ towards $\lambda_{k}(\Delta_P)$.
@@
 
-In the case where $\mc{X}$ is assumed to have a boundary, the graph Laplacian $L_{n,\varepsilon}$ is a reasonable approximation of the operator $\Delta_P$ at points $x \in \mc{X}$ for which $B(x,\varepsilon) \subseteq \mc{X}$.
+In the case where $\mc{X}$ is assumed to have a boundary, the graph Laplacian $L_{n,\varepsilon}$ is a reasonable approximation of the operator $\Delta_P$ only at points $x \in \mc{X}$ for which $B(x,\varepsilon) \subseteq \mc{X}$.
 In contrast, at points $x$ near the boundary of $\mc{X}$, the graph Laplacian is known to approximate a different operator altogether \citep{belkin2012}.\footnote{This is directly related to the boundary bias of kernel smoothing, since the graph Laplacian can be viewed as a kernel-based estimator of $\Delta_P$.} This renders analysis of $\lambda_k$ substantially more challenging, since its continuum limit is not $\lambda_k(\Delta_P)$.
@@
 
-\paragraph{Empirical norm.}
+\subsubsection{Empirical norm}
 Finally, in Proposition~\ref{prop:empirical_norm_sobolev} we establish that a one-sided bound of the form $\|f_0\|_n^2 \gtrsim \|f_0\|_P^2$ whenever $\|f_0\|_P^2$ is itself sufficiently large.
@@
 	\label{prop:empirical_norm_sobolev}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f \in H^s(\Xset,M)$ for some $s > d/4$.
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f \in H^s(\mc{X},M)$ for some $s > d/4$.
 There exist constants $c$ and $C$ that do not depend on $f_0$ or $n$ such that the following statement holds for any $\delta > 0$:  if
@@
 	\label{eqn:empirical_norm_sobolev_1}
-	\norm{f}_{P} \geq C M \biggl(\frac{1}{\delta n}\biggr)^{s/d}
+	\|f\|_{P} \geq C M \biggl(\frac{1}{\delta n}\biggr)^{s/d}
 	\end{equation}
@@
 	\label{eqn:empirical_norm_sobolev}
-	\norm{f}_n^2 \geq \frac{1}{2} \|f_0\|_P^2.
+	\|f\|_n^2 \geq \frac{1}{2} \|f_0\|_P^2.
 	\end{equation}
@@
 In this section we consider the manifold setting, where $(X_1,Y_1),\ldots,(X_n,Y_n)$ are observed according to Model~\ref{def:model_manifold}.
-In this setting, it is known that the minimax rates depend only on the intrinsic dimension $m$; more specifically,~\cite{bickel2007,ariascastro2018} show that for functions with H\"{o}lder smoothness $s$, the minimax estimation rate is $n^{-2s/(2s + m)}$ and the testing rate is $n^{-4s/(4s + m)}$.\footnote{Although~\cite{ariascastro2018} considers density testing, usual arguments regarding equivalence of experiments~\citep{brown96} imply that the same rates apply to regression testing.} On the other hand, a theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
+In this setting, it is known that the minimax rates depend only on the intrinsic dimension $m$; more specifically,~\cite{bickel2007,ariascastro2018} show that for functions with H\"{o}lder smoothness $s$, the minimax estimation rate is $n^{-2s/(2s + m)}$ and the testing rate is $n^{-4s/(4s + m)}$.\footnote{Although~\cite{ariascastro2018} considers density testing, usual arguments regarding equivalence of experiments~\citep{brown96} imply that the same rates apply to regression testing.} On the other hand, a theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G$ can ``learn'' the manifold $\mc{X}$ in various senses, so long as $\mc{X}$ is locally linear.
 We build on this work by showing that when $P$ is supported on a manifold $\mc{X}$ and $f_0 \in H^s(\mc{X})$, Laplacian eigenmaps achieves the sharper minimax estimation and testing rates.
@@
 
-\paragraph{Estimation.}
+\subsubsection{Estimation}
 To ensure that $\wh{f}$ is an in-sample minimax rate-optimal estimator, we choose the kernel function $\eta$, graph radius $\varepsilon$ and number of eigenvectors $K$ as in~\ref{asmp:parameters_estimation_ho}, except with ambient dimension $d$ replaced by the intrinsic dimension $m$.
@@
 
-\paragraph{Testing.}
+\subsubsection{Testing}
 Likewise, to construct a minimax optimal test using $\wh{T}$, we choose $\varepsilon$ and $K$ as in~\ref{asmp:parameters_testing_fo}, except with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
@@
 
-\paragraph{Analysis.}
+\subsubsection{Analysis}
 The high-level strategy used to prove Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold} is the same as in the flat-Euclidean setting.
@@
 	\label{prop:empirical_norm_sobolev_manifold}
-	Suppose Model~\ref{def:model_manifold}, and additionally that $f_0 \in H^s(\Xset,M)$ for some $s > m/4$.
+	Suppose Model~\ref{def:model_manifold}, and additionally that $f_0 \in H^s(\mc{X},M)$ for some $s > m/4$.
 There exists a constant $C$ that does not depend on $f_0$ such that the following statement holds for all $\delta > 0$:  if
@@
 	\label{eqn:empirical_norm_sobolev_manifold_1}
-	\norm{f_0}_{P} \geq \frac{C M}{\delta^{s/m}}n^{-s/m},
+	\|f_0\|_{P} \geq \frac{C M}{\delta^{s/m}}n^{-s/m},
 	\end{equation}
@@
 	\label{eqn:empirical_norm_sobolev_manifold}
-	\norm{f_0}_n^2 \geq \frac{1}{2} \|f_0\|_P^2.
+	\|f_0\|_n^2 \geq \frac{1}{2} \|f_0\|_P^2.
 	\end{equation}
@@
 \label{sec:out_of_sample}
-Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity} show that $\wh{f}$ is a minimax optimal estimator over Sobolev spaces, when the error is measured \emph{in-sample}---that is, in $\Leb^2(P_n)$ norm.
+Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity} show that $\wh{f}$ is a minimax optimal estimator over Sobolev spaces, when the error is measured \emph{in-sample}---that is, in $L^2(P_n)$ norm.
 In this section we consider \emph{out-of-sample} error---measured in $L^2(P)$ norm---which is the more typical metric used in the study of nonparametric regression with random design.
 
-Of course, the Laplacian eigenmaps estimator is only defined at the design points $\{X_1,\ldots,X_n\}$, and to measure error in $L^2(P)$ norm we must first extend it to the entire domain $\Xset$.
+Of course, the Laplacian eigenmaps estimator is only defined at the design points $\{X_1,\ldots,X_n\}$, and to measure error in $L^2(P)$ norm we must first extend it to the entire domain $\mc{X}$.
 We propose a simple method, kernel smoothing, to do the job.
@@
 
-\paragraph{Extension by kernel smoothing.}
+\subsection{Extension by kernel smoothing}
 We now formally define our approach to extension by kernel smoothing.
@@
 
-\paragraph{Out-of-sample error of kernel smoothed Laplacian eigenmaps.}
+\subsection{Out-of-sample error of kernel smoothed Laplacian eigenmaps}
 
@@
 Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data, by a factor of $h^2$.
-On the other hand the bias term is typical, and when $\psi$ is an order-$s$ kernel a standard analysis (see Lemma~\ref{lem:kernel_smoothing_bias}) shows that $\|T_{P,h}f_0 - f_0\|_P^2 \lesssim |f|_{H^s(\mc{X})}^2 h^{2s}$.
+On the other hand the bias term is typical, and when $\psi$ is an order-$s$ kernel a standard analysis (see Lemma~16 in the Supplementary Material) shows that $\|T_{P,h}f_0 - f_0\|_P^2 \lesssim |f|_{H^s(\mc{X})}^2 h^{2s}$.
 \begin{enumerate}[label=(K\arabic*)]
@@
 \begin{figure*}[tb]
-	\includegraphics[width=.245\textwidth]{figures/mse/mse_by_sample_size_1d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/mse/mse_by_sample_size_1d_2s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/mse/mse_by_sample_size_2d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/mse/mse_by_sample_size_2d_2s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_1d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_1d_2s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_2d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_2d_2s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/mse/mse_by_sample_size_1d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/mse/mse_by_sample_size_1d_2s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/mse/mse_by_sample_size_2d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/mse/mse_by_sample_size_2d_2s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/out_of_sample_mse/mse_by_sample_size_1d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/out_of_sample_mse/mse_by_sample_size_1d_2s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/out_of_sample_mse/mse_by_sample_size_2d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/out_of_sample_mse/mse_by_sample_size_2d_2s.pdf}
 	\caption{Mean squared error (mse) of Laplacian eigenmaps and spectral projection estimators.
@@
 
-\paragraph{Estimation.}
+\subsection{Estimation}
 In our first experiment, we compare the mean-squared error of the Laplacian eigenmaps estimator $\wh{f}$  to that of its population-level counterpart $\wt{f}$.
@@
 
-\paragraph{Testing.} 
+\subsection{Testing} 
 In our second experiment, we compare the Laplacian eigenmaps test $\varphi$ against the  population-level spectral projection test $\wt{\varphi}$.
@@
 \begin{figure*}[tb]
-	\includegraphics[width=.245\textwidth]{figures/testing/eigenfunction/critical_radius_by_sample_size_1d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/testing/eigenfunction/critical_radius_by_sample_size_1d_2s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/testing/eigenfunction/critical_radius_by_sample_size_2d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/testing/eigenfunction/critical_radius_by_sample_size_2d_2s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/testing/eigenfunction/critical_radius_by_sample_size_1d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/testing/eigenfunction/critical_radius_by_sample_size_1d_2s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/testing/eigenfunction/critical_radius_by_sample_size_2d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/testing/eigenfunction/critical_radius_by_sample_size_2d_2s.pdf}
 	\caption{Worst-case testing risk Laplacian eigenmaps (\texttt{LE}) and spectral projection (\texttt{SP}) tests, as a function of sample size $n$.
@@
 
-\paragraph{Tuning parameters.}
+\subsection{Tuning parameters}
 Our first two experiments demonstrate that Laplacian eigenmaps methods have comparable statistical performance to spectral projection methods.
@@
 \begin{figure*}[tb]
-	\includegraphics[width=.245\textwidth]{figures/tuning/eigenfunction/mse_by_number_of_eigenvectors_1d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/tuning/eigenfunction/mse_by_radius_1d_1s.pdf} 
-	\includegraphics[width=.245\textwidth]{figures/tuning/eigenfunction/mse_by_number_of_eigenvectors_2d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/tuning/eigenfunction/mse_by_radius_2d_1s.pdf} 
-	\includegraphics[width=.245\textwidth]{figures/tuning/sobolev/mse_by_number_of_eigenvectors_1d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/tuning/sobolev/mse_by_radius_1d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/tuning/sobolev/mse_by_number_of_eigenvectors_2d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/tuning/sobolev/mse_by_radius_2d_1s.pdf}  
+	\includegraphics[width=.245\textwidth]{../figures/tuning/eigenfunction/mse_by_number_of_eigenvectors_1d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/tuning/eigenfunction/mse_by_radius_1d_1s.pdf} 
+	\includegraphics[width=.245\textwidth]{../figures/tuning/eigenfunction/mse_by_number_of_eigenvectors_2d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/tuning/eigenfunction/mse_by_radius_2d_1s.pdf} 
+	\includegraphics[width=.245\textwidth]{../figures/tuning/sobolev/mse_by_number_of_eigenvectors_1d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/tuning/sobolev/mse_by_radius_1d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/tuning/sobolev/mse_by_number_of_eigenvectors_2d_1s.pdf}
+	\includegraphics[width=.245\textwidth]{../figures/tuning/sobolev/mse_by_radius_2d_1s.pdf}  
 	\caption{Mean squared error of Laplacian eigenmaps (\textcolor{red}{red}), Laplacian eigenmaps plus kernel smoothing (\textcolor{blue}{blue}), and spectral projection (\textcolor{green}{green}) as a function of tuning parameters.
@@
 using the estimator $T_{n,h}{\bf Y}$ (possibly for a different choice of $h$).
-In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{n,h}\wh{f}\|_P^2$ is of a strictly lower order than $\inf_{h'} \Ebb\|f_0 - T_{n,h'}{\bf Y}\|_P^2$.
+In Appendix~F.4 of the Supplementary Material, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{n,h}\wh{f}\|_P^2$ is of a strictly lower order than $\inf_{h'} \Ebb\|f_0 - T_{n,h'}{\bf Y}\|_P^2$.
 This is possible because Laplacian eigenmaps induces a completely different bias than kernel smoothing.
@@
 
-\bibliographystyle{plainnat}
-\bibliography{../../graph_regression_bibliography}
-
-\appendix
-
-\input{appendix.tex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Support information, if any,             %%
+%% should be provided in the                %%
+%% Acknowledgements section.                %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\begin{acks}[Acknowledgments]
+%
+%\end{acks}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Funding information, if any,             %%
+%% should be provided in the                %%
+%% funding section.                         %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\begin{funding}
+%\end{funding}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Supplementary Material, including data   %%
+%% sets and code, should be provided in     %%
+%% {supplement} environment with title      %%
+%% and short description. It cannot be      %%
+%% available exclusively as external link.  %%
+%% All Supplementary Material must be       %%
+%% available to the reader on Project       %%
+%% Euclid with the published article.       %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{supplement}
+\stitle{Supplement to ``Minimax-Optimal Regression Over Sobolev Spaces via Laplacian Eigenmaps with Neighborhood Graphs''.}
+\sdescription{The supplement contains all proofs omitted from the main text due to space constraints.}
+\end{supplement}
+
+% Bibliography
+\bibliographystyle{imsart-nameyear}
+\bibliography{../../../graph_regression_bibliography}
 

### 2021-09-04 manuscripts/laplacian_eigenmaps/annals_of_statistics/supplement.tex
Put Laplacian Eigenmaps paper into AOS format.

--- 2021-08-31 manuscripts/laplacian_eigenmaps/appendix.tex
@@
-\noindent
+\documentclass[aos]{imsart}
+
+%% Packages
+\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
+% \RequirePackage[numbers]{natbib}
+\RequirePackage[authoryear]{natbib}
+\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
+\RequirePackage{graphicx}
+\usepackage{mathtools, enumitem}
+\usepackage{multirow}
+
+\startlocaldefs
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% For Axiom, Claim, Corollary, Hypothesis, %%
+%% Lemma, Theorem, Proposition              %%
+%% use \theoremstyle{plain}                 %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
+\newtheorem{corollary}{Corollary}
+\newtheorem{proposition}{Proposition}
+\theoremstyle{definition}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% For Assumption, Definition, Example,     %%
+%% Notation, Property, Remark, Fact         %%
+%% use \theoremstyle{remark}                %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\theoremstyle{remark}
+\newtheorem{remark}{Remark}
+\newtheorem{definition}{Definition}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Please put your definitions here:        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%% Begin Ryan's definitions
+
+% Widebar
+\makeatletter
+\newcommand*\rel@kern[1]{\kern#1\dimexpr\macc@kerna}
+\newcommand*\widebar[1]{%
+	\begingroup
+	\def\mathaccent##1##2{%
+		\rel@kern{0.8}%
+		\overline{\rel@kern{-0.8}\macc@nucleus\rel@kern{0.2}}%
+		\rel@kern{-0.2}%
+	}%
+	\macc@depth\@ne
+	\let\math@bgroup\@empty \let\math@egroup\macc@set@skewchar
+	\mathsurround\z@ \frozen@everymath{\mathgroup\macc@group\relax}%
+	\macc@set@skewchar\relax
+	\let\mathaccentV\macc@nested@a
+	\macc@nested@a\relax111{#1}%
+	\endgroup
+}
+\makeatother
+
+% Min and max
+\newcommand{\argmin}{\mathop{\mathrm{argmin}}}
+\newcommand{\argmax}{\mathop{\mathrm{argmax}}}
+\newcommand{\minimize}{\mathop{\mathrm{minimize}}}
+\newcommand{\st}{\mathop{\mathrm{subject\,\,to}}}
+
+% Shortcuts
+\def\R{\mathbb{R}}
+
+%%% End Ryan's definitions
+
+%%% Begin Alden's additions
+\newcommand{\Ebb}{\mathbb{E}}
+\newcommand{\Pbb}{\mathbb{P}}
+\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
+\newcommand{\wt}[1]{\widetilde{#1}}
+\newcommand{\wh}[1]{\widehat{#1}}
+\newcommand{\mc}[1]{\mathcal{#1}}
+\newcommand{\Reals}{\mathbb{R}} % Same thing as Ryan's \R
+\newcommand{\Rd}{\Reals^d}
+\newcommand{\wb}[1]{\widebar{#1}}
+\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
+\newcommand{\Var}{\mathrm{Var}}
+\newcommand{\Cov}{\mathrm{Cov}}
+\newcommand{\1}{\mathbf{1}}
+\newcommand{\bj}{{\bf j}}
+\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
+
+\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
+\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
+\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
+\DeclareMathAccent{\wc}{0}{mathx}{"71}
+%%% End Alden's Additions
+
+\endlocaldefs
+
+\begin{document}
+
+\begin{frontmatter}
+
+\title{Supplementary Material For: Minimax-optimal Regression over Sobolev Spaces via Laplacian Eigenmaps with Neighborhood Graphs}
+%\thankstext{T1}{A sample of additional note to the title.}
+
+\begin{aug}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Only one address is permitted per author. %%
+%% Only division, organization and e-mail is %%
+%% included in the address.                  %%
+%% Additional information can be included in %%
+%% the Acknowledgments section if necessary. %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\author[A]{\fnms{Alden} \snm{Green}\ead[label=e1]{}},
+\author[A]{\fnms{Sivaraman} \snm{Balakrishnan}\ead[label=e1]{}}
+\and
+\author[A]{\fnms{Ryan J.} \snm{Tibshirani}\ead[label=e1]{}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Addresses                                %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\address[A]{Carnegie Mellon University, \printead{e1}}
+\end{aug}
+
+\end{frontmatter}
+% Main text
 
@@
 	\label{lem:fixed_graph_estimation}
-	For any integer $s > 0$, and any integer $0 \leq K \leq n$, the Laplacian eigenmaps estimator $\wh{f}$ of~\eqref{eqn:laplacian_eigenmaps_estimator} satisfies
+	For any integer $s > 0$, and any integer $0 \leq K \leq n$, the Laplacian eigenmaps estimator $\wh{f}$ satisfies
 	\begin{equation}
@@
 \label{sec:graph_quadratic_form_euclidean}
-In this section we prove Proposition~\ref{prop:graph_seminorm_ho}. The proposition will follow from several intermediate results.
+% AG: Hard-coded reference.
+In this section we prove Proposition~4. The proposition will follow from several intermediate results.
 \begin{enumerate}
@@
 
-\paragraph{Proof (of Proposition~\ref{prop:graph_seminorm_ho}).}
-Proposition~\ref{prop:graph_seminorm_ho} follows immediately from~\eqref{pf:graph_seminorm_ho_1} and~\eqref{pf:graph_seminorm_ho_2}. \qed
+\paragraph{Proof (of Proposition~4).}
+Proposition~4 follows immediately from~\eqref{pf:graph_seminorm_ho_1} and~\eqref{pf:graph_seminorm_ho_2}. \qed
 
@@
 	\label{lem:approximation_error_nonlocal_laplacian}
-	Assume Model~\ref{def:model_flat_euclidean}.
+	Assume Model~1.
 Let $s \in \mathbb{N} \setminus \{0,1\}$, suppose that $f \in H^s(\mc{X};M)$, and if $s > 1$ suppose that $p \in C^{s - 1}(\mc{X})$.
-Let $L_{P,\varepsilon}$ be define with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_flat_euclidean}.
+Let $L_{P,\varepsilon}$ be define with respect to a kernel $\eta$ that satisfies~(K1).
 Then there exist constants $C_1$ and $C_2$ that do not depend on $f$, such that each of the following statements hold.
@@
 \end{equation*}
-for some $g_s \in \Leb^2(\mc{X}_{k\varepsilon},C \cdot M \cdot \|p\|_{C^{s - 1}(\mc{X})})$, and integrating over $\mc{X}_{k\varepsilon}$ gives the desired result.
+for some $g_s \in L^2(\mc{X}_{k\varepsilon},C \cdot M \cdot \|p\|_{C^{s - 1}(\mc{X})})$, and integrating over $\mc{X}_{k\varepsilon}$ gives the desired result.
 \eqref{eqn:approximation_error_nonlocal_laplacian_2} follows from~\eqref{pf:approximation_error_nonlocal_laplacian_1} in an identical fashion.
@@
 \label{pf:approximation_error_nonlocal_laplacian_2}
-\|G_2\|_{\Leb^2(\mc{X}_{\varepsilon})} \leq C \varepsilon^{s - 2} M \|p\|_{C^{s - 1}(\mc{X})};
+\|G_2\|_{L^2(\mc{X}_{\varepsilon})} \leq C \varepsilon^{s - 2} M \|p\|_{C^{s - 1}(\mc{X})};
 \end{equation}
@@
 \end{align}
-Furthermore, for each $j = 1,\ldots,s - 1$ convolution of $d_x^jf$ with $\eta$ only decreases the $\Leb^2(\mc{X}_{\varepsilon})$ norm, meaning
+Furthermore, for each $j = 1,\ldots,s - 1$ convolution of $d_x^jf$ with $\eta$ only decreases the $L^2(\mc{X}_{\varepsilon})$ norm, meaning
 \begin{equation}
@@
 & \leq \int_{B(0,1)} \int_{\mc{X}_{\varepsilon}} \Bigl[\bigl(d^jf\bigr)(x)\Bigr]^2 \eta(\|z\|) \,dx  \,dz \\
-& \leq \|d^jf\|_{\Leb^2(\mc{X_{\varepsilon}})}^2.
+& \leq \|d^jf\|_{L^2(\mc{X_{\varepsilon}})}^2.
 \end{aligned}
@@
 \int_{\mc{X}_{\varepsilon}} |G_2(x)|^2 \,dx & \leq C \Bigl(\varepsilon^{s - 2}\|p\|_{C^{s - 1}(\mc{X})}\Bigr)^2 \sum_{j = 1}^{s - 1} \int_{\mc{X}_{\varepsilon}}\biggl(\frac{1}{j!} \int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(z)\Bigr| \cdot \Bigl|\eta(\|z\|)\Bigr| \,dz\biggr)^2 \,dx \\
-& \leq C \Bigl(\varepsilon^{s - 2}\|p\|_{C^{s - 1}(\mc{X})}\Bigr)^2 \sum_{j = 1}^{s - 1} \|d^ju\|_{\Leb^2(\mc{X_{\varepsilon}})}^2,
+& \leq C \Bigl(\varepsilon^{s - 2}\|p\|_{C^{s - 1}(\mc{X})}\Bigr)^2 \sum_{j = 1}^{s - 1} \|d^ju\|_{L^2(\mc{X_{\varepsilon}})}^2,
 \end{align*}
@@
 with the last inequality following from~\eqref{eqn:sobolev_remainder_term}.
-Noting that $p_{\max} = \|p\|_{C^0(\mc{X})} \leq \|p\|_{C^{s - 1}(\mc{X})}$, we see that this is a sufficient bound on $\|G_3\|_{\Leb^2(\mc{X}_{\varepsilon})}$.
+Noting that $p_{\max} = \|p\|_{C^0(\mc{X})} \leq \|p\|_{C^{s - 1}(\mc{X})}$, we see that this is a sufficient bound on $\|G_3\|_{L^2(\mc{X}_{\varepsilon})}$.
 
@@
 
-In Lemma~\ref{lem:approximation_error_nonlocal_laplacian_boundary}, we establish that if $f$ is Sobolev smooth of order $s > 2k$ and zero-trace, then near the boundary of $\mc{X}$ the non-local Laplacian $L_{P,\varepsilon}^kf$ is close to $0$ in the $\Leb^2$-sense.
+In Lemma~\ref{lem:approximation_error_nonlocal_laplacian_boundary}, we establish that if $f$ is Sobolev smooth of order $s > 2k$ and zero-trace, then near the boundary of $\mc{X}$ the non-local Laplacian $L_{P,\varepsilon}^kf$ is close to $0$ in the $L^2$-sense.
 \begin{lemma}
 	\label{lem:approximation_error_nonlocal_laplacian_boundary}
-	Assume Model~\ref{def:model_flat_euclidean}.
-Let $s,k \in \mathbb{N}$.
-Suppose that $f \in H_0^{s}(\mc{X};M)$.
-Then there exist numbers $c,C > 0$ that do not depend on $M$, such that for all $\varepsilon < c$,
+	Assume Model~1. Let $s,k \in \mathbb{N}$. Suppose that $f \in H_0^{s}(\mc{X};M)$. Then there exist numbers $c,C > 0$ that do not depend on $M$, such that for all $\varepsilon < c$, 
 	\begin{equation*}
@@
 Noticing that $\Delta_P^kf \in H^1(\mc{X};\|p\|_{C^{s - 1}(\mc{X})}^kM)$, we use~\eqref{pf:estimate_nonlocal_seminorm_1} to conclude that $E_{P,\varepsilon}(\Delta_P^kf;\mc{X}_{k\varepsilon}) \leq C_0M^2$.
-On the other hand, it follows from Assumption~\ref{asmp:kernel_flat_euclidean} and~\eqref{eqn:approximation_error_nonlocal_laplacian_1} that
+On the other hand, it follows from Assumption~(K1) and~\eqref{eqn:approximation_error_nonlocal_laplacian_1} that
 \begin{align*}
@@
 \label{pf:estimate_nonlocal_seminorm_3}
-\|L_{P,\varepsilon}^{k + 1}f\|_P^2 \leq p_{\max} \|L_{P,\varepsilon}^{k + 1}f\|_{\Leb^2(\mc{X})}^2 \leq p_{\max}\biggl\{\int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^{k + 1}f(x)\bigr)^2 \,dP(x) + \int_{\partial_{k\varepsilon}\mc{X}} \bigl(L_{P,\varepsilon}^{k + 1}f(x)\bigr)^2 \,dP(x)\biggr\},
+\|L_{P,\varepsilon}^{k + 1}f\|_P^2 \leq p_{\max} \|L_{P,\varepsilon}^{k + 1}f\|_{L^2(\mc{X})}^2 \leq p_{\max}\biggl\{\int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^{k + 1}f(x)\bigr)^2 \,dP(x) + \int_{\partial_{k\varepsilon}\mc{X}} \bigl(L_{P,\varepsilon}^{k + 1}f(x)\bigr)^2 \,dP(x)\biggr\},
 \end{equation}
@@
 	\label{lem:l2estimate_nonlocal_laplacian}
-	Assume Model~\ref{def:model_flat_euclidean}.
-Suppose $f \in L^2(U;M)$ for a Borel set $U \subseteq \mc{X}$, and let $L_{P,\varepsilon}$ be defined with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_flat_euclidean}.
+	Assume Model~1.
+Suppose $f \in L^2(U;M)$ for a Borel set $U \subseteq \mc{X}$, and let $L_{P,\varepsilon}$ be defined with respect to a kernel $\eta$ that satisfies~(K1).
 Then there exists a constant $C$ which does not depend on $f$ or $M$ such that
@@
 	\label{eqn:l2estimate_nonlocal_laplacian}
-	\|L_{P,\varepsilon}f\|_{\Leb^2(U)} \leq \frac{2 p_{\max}}{\varepsilon^2} \|f\|_{L^2(U)}
+	\|L_{P,\varepsilon}f\|_{L^2(U)} \leq \frac{2 p_{\max}}{\varepsilon^2} \|f\|_{L^2(U)}
 	\end{equation}
@@
 	\label{lem:dirichlet_estimate_nonlocal_laplacian}
-	Assume Model~\ref{def:model_flat_euclidean}.
-Suppose $f \in L^2(U;M)$ for a Borel set $U \subseteq \mc{X}$, and let $L_{P,\varepsilon}$ be defined with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_flat_euclidean}.
+	Assume Model~1.
+Suppose $f \in L^2(U;M)$ for a Borel set $U \subseteq \mc{X}$, and let $L_{P,\varepsilon}$ be defined with respect to a kernel $\eta$ that satisfies~(K1).
 Then there exists a constant $C$ which does not depend on $f$ or $M$ such that
@@
 	\label{lem:graph_seminorm_bias2}
-	Assume Model~\ref{def:model_flat_euclidean}.
-Suppose $f \in H^1(\mc{X};M)$, and let $D_if$ be defined with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_flat_euclidean}.
+	Assume Model~1.
+Suppose $f \in H^1(\mc{X};M)$, and let $D_if$ be defined with respect to a kernel $\eta$ that satisfies~(K1).
 Then there exists a constant $C$ which does not depend on $f$ or $M$, such that for any $i \in [n]$ and $\bj \in [n]^s$,
@@
 \paragraph{Proof (of Lemma~\ref{lem:l2estimate_nonlocal_laplacian}).}
-We fix a version of $f \in \Leb^2(U)$, so that we may speak of its pointwise values.
+We fix a version of $f \in L^2(U)$, so that we may speak of its pointwise values.
 
@@
 \end{align*}
-The equality follows by the assumption $\int_{\Rd} \eta(\|z\|)\,dx = 1$ in~\ref{asmp:kernel_flat_euclidean}. Integrating over all $x \in U$, it follows from the triangle inequality that
+The equality follows by the assumption $\int_{\Rd} \eta(\|z\|)\,dx = 1$ in~(K1). Integrating over all $x \in U$, it follows from the triangle inequality that
 \begin{align}
-\|L_{P,\varepsilon}\|_{\Leb^2(U)}^2 & \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx \nonumber \\
+\|L_{P,\varepsilon}\|_{L^2(U)}^2 & \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx \nonumber \\
 & \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx.
@@
 \begin{equation}
-\int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx = 2 \int_{U} \int_{U} |f(x)|^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx \leq 2 \varepsilon^d \int_{U} |f(x)|^2 \,dx = 2\varepsilon^d \|f\|_{\Leb^2(U)}^2,
+\int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx = 2 \int_{U} \int_{U} |f(x)|^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx \leq 2 \varepsilon^d \int_{U} |f(x)|^2 \,dx = 2\varepsilon^d \|f\|_{L^2(U)}^2,
 \label{pf:l2estimate_nonlocal_laplacian_2}
@@
 \begin{equation*}
-\|L_{P,\varepsilon}\|_{\Leb^2(U)}^2 \leq \frac{4p_{\max}^2}{\varepsilon^4} \|f\|_{\Leb^2(U)}^2. 
+\|L_{P,\varepsilon}\|_{L^2(U)}^2 \leq \frac{4p_{\max}^2}{\varepsilon^4} \|f\|_{L^2(U)}^2. 
 \end{equation*}
@@
 \label{sec:graph_quadratic_form_manifold}
-In this section we prove Proposition~\ref{prop:graph_seminorm_manifold}.
-Note that when $s = 1$, the upper bound~\eqref{eqn:graph_seminorm_manifold} follows immediately from Lemma~\ref{lem:dirichlet_energy_sobolev} and Markov's inequality.
-
-On the other hand when $s = 2$ or $s = 3$, we prove Proposition~\ref{prop:graph_seminorm_manifold} by first establishing some intermediate results, many of which are analogous to results we have already shown in the flat Euclidean case.
+In this section we prove Proposition~7.
+Note that when $s = 1$, the desired upper bound (Equation~(42) in the main text) follows immediately from Lemma~\ref{lem:dirichlet_energy_sobolev} and Markov's inequality.
+
+On the other hand when $s = 2$ or $s = 3$, we prove Proposition~7 by first establishing some intermediate results, many of which are analogous to results we have already shown in the flat Euclidean case.
 Indeed, in some ways the proof will be simpler in the manifold setting than in the flat Euclidean case: there is no boundary, and we do not need to analyze the iterated nonlocal Laplacian $L_{P,\varepsilon}^j$ for $j > 1$.
@@
 
-We show the following results, each of which hold under the same assumptions as Proposition~\ref{prop:graph_seminorm_manifold}.
+We show the following results, each of which hold under the same assumptions as Proposition~7.
 \begin{itemize}
@@
 	\label{eqn:nonlocal_laplacian_approximation_error_manifold_l2}
-	\|\wt{L}_{P,\varepsilon}f\|_{\Leb^2(\mc{X})}^2 \leq C_3 \|f\|_{H^2(\mc{X})}^2,
+	\|\wt{L}_{P,\varepsilon}f\|_{L^2(\mc{X})}^2 \leq C_3 \|f\|_{H^2(\mc{X})}^2,
 	\end{equation}
@@
 	\label{eqn:nonlocal_laplacian_approximation_error_manifold_sobolev}
-	\|\wt{L}_{P,\varepsilon}f - \sigma_{\eta}\Delta_Pf\|_{\Leb^2(\mc{X})}^2 \leq C_3 \varepsilon^2 \|f\|_{H^3(\mc{X})}^2.
+	\|\wt{L}_{P,\varepsilon}f - \sigma_{\eta}\Delta_Pf\|_{L^2(\mc{X})}^2 \leq C_3 \varepsilon^2 \|f\|_{H^3(\mc{X})}^2.
 	\end{equation}
@@
 
-\paragraph{Proof (of Proposition~\ref{prop:graph_seminorm_manifold}).} Follows immediately from~\eqref{pf:graph_seminorm_manifold_1} and~\eqref{eqn:manifold_nonlocal_seminorm}. \qed
+\paragraph{Proof (of Proposition~7).} Follows immediately from~\eqref{pf:graph_seminorm_manifold_1} and~\eqref{eqn:manifold_nonlocal_seminorm}. \qed
 
@@
 \begin{equation*}
-\|A_1\|_{\Leb^2(\mc{X})}^2 \leq C \|f\|_{H^2(\mc{X})}^2
+\|A_1\|_{L^2(\mc{X})}^2 \leq C \|f\|_{H^2(\mc{X})}^2
 \end{equation*}
@@
 \begin{equation*}
-\|A_1 - \sigma_{\eta} \Delta_Pf\|_{\Leb^2(\mc{X})}^2 \leq C\|f\|_{H^3(\mc{X})}^2 \varepsilon^2.
-\end{equation*}
-
-Therefore it remains only to upper bound $A_2$ in $\Leb^2(\mc{X})$ norm.
+\|A_1 - \sigma_{\eta} \Delta_Pf\|_{L^2(\mc{X})}^2 \leq C\|f\|_{H^3(\mc{X})}^2 \varepsilon^2.
+\end{equation*}
+
+Therefore it remains only to upper bound $A_2$ in $L^2(\mc{X})$ norm.
 To do so, we recall (1.34) of \cite{trillos2019}: for any $\varepsilon < i_0$ and all $x \in \mc{X}$, the Jacobian $J_x(v)$ satisfies the upper bound
@@
 \begin{align*}
-\|A_2\|_{\Leb^2(\mc{X})}^2 & \leq Cm^2K_0^2 \biggl[\int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr)^2 \eta(\|v\|) q(\varepsilon v) \,dv\biggr] \cdot \biggl[\int_{B(0,1)} \eta(\|v\|) q(\varepsilon v) \,dv\biggr] \\
+\|A_2\|_{L^2(\mc{X})}^2 & \leq Cm^2K_0^2 \biggl[\int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr)^2 \eta(\|v\|) q(\varepsilon v) \,dv\biggr] \cdot \biggl[\int_{B(0,1)} \eta(\|v\|) q(\varepsilon v) \,dv\biggr] \\
 & \leq Cm^2K_0^2 \sigma_{\eta} (1 + L_q\varepsilon)  \int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr)^2 \eta(\|v\|) q(\varepsilon v) \,dv \\
@@
 \begin{equation*}
- E_{P,\varepsilon}(L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f;\mc{X}) \leq  \frac{C}{\varepsilon^2}\|L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f\|_{\Leb^2(\mc{X})}^2 \leq C |f|_{H^1(\mc{X})}^2.
+E_{P,\varepsilon}(L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f;\mc{X}) \leq  \frac{C}{\varepsilon^2}\|L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f\|_{L^2(\mc{X})}^2 \leq C |f|_{H^1(\mc{X})}^2.
 \end{equation*}
@@
 Recall the Dirichlet energy $E_{P,\varepsilon}(f;\mc{X}) = \dotp{L_{P,\varepsilon}f}{f}_P$, defined in~\eqref{eqn:dirichlet_energy}.
-Now we establish some estimates on $E_{P,\varepsilon}(f;\mc{X})$ under Model~\ref{def:model_manifold}, and under various assumptions regarding the regularity of $f$.
+Now we establish some estimates on $E_{P,\varepsilon}(f;\mc{X})$ under Model~2, and under various assumptions regarding the regularity of $f$.
 \begin{lemma}
 	\label{lem:dirichlet_energy_l2}
-	Suppose Model~\ref{def:model_manifold}, and additionally that $f \in L^2(\mc{X})$. Then there exists a constant $C$ such that
+	Suppose Model~2, and additionally that $f \in L^2(\mc{X})$. Then there exists a constant $C$ such that
 	\begin{equation}
@@
 	\label{lem:dirichlet_energy_sobolev}
-	Suppose Model~\ref{def:model_manifold}, and additionally that $f \in H^1(\mc{X})$. Then there exist constants $c$ and $C$ which do not depend on $f$ such that for any $0 < \varepsilon < c$,
+	Suppose Model~2, and additionally that $f \in H^1(\mc{X})$. Then there exist constants $c$ and $C$ which do not depend on $f$ such that for any $0 < \varepsilon < c$,
 	\begin{equation}
@@
 	\label{lem:dirichlet_energy_remainder}
-	Suppose Model~\ref{def:model_manifold}, and additionally that $f \in H^1(\mc{X})$. There exist constants $c$ and $C$ such that for any $\varepsilon < c$,
+	Suppose Model~2, and additionally that $f \in H^1(\mc{X})$. There exist constants $c$ and $C$ such that for any $\varepsilon < c$,
 	\begin{equation}
@@
 Finally, we use Lemma~\ref{lem:manifold_graph_seminorm_bias2} to show that the pure bias component of $\dotp{L_n^sf,f}_n$ is small in expectation.
-This is analogous to Lemma~\ref{lem:graph_seminorm_bias2}, except assuming Model~\ref{def:model_manifold} rather than Model~\ref{def:model_flat_euclidean}.
+This is analogous to Lemma~\ref{lem:graph_seminorm_bias2}, except assuming Model~2 rather than Model~1.
 \begin{lemma}
 	\label{lem:manifold_graph_seminorm_bias2}
-	Assume Model~\ref{def:model_manifold}.
-Suppose $f \in H^1(\mc{X})$, and let $D_if$ be defined with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_manifold}.
+	Assume Model~2.
+Suppose $f \in H^1(\mc{X})$, and let $D_if$ be defined with respect to a kernel $\eta$ that satisfies~(P5).
 Then there exists a constant $C$ which does not depend on $f$ or $n$, such that for any $i \in [n]$ and $\bj \in [n]^s$,
@@
 \begin{equation*}
-\int_{\mc{X}} \int_{B_{m}(0,\wt{\varepsilon})} |d_x^{1}f(v)|^2 \,dv \,d\mu(x)  = \frac{\nu_m}{2 + m} (\wt{\varepsilon}^{2 + m} - \varepsilon^{2 + m}) \int_{\mc{X}} |d_x^1f|^2 \,d\mu(x) = 27\frac{\nu_m}{(2 + m)R^2} \varepsilon^{4 + m} \|d^1f\|_{\Leb^2(\mc{X})}^2.
-\end{equation*}
-Recalling that $\|d^1f\|_{\Leb^2(\mc{X})}^2 \leq \|f\|_{H^1(\mc{X})}^2$, we see that this implies the claim of Lemma~\ref{lem:dirichlet_energy_remainder}.
+\int_{\mc{X}} \int_{B_{m}(0,\wt{\varepsilon})} |d_x^{1}f(v)|^2 \,dv \,d\mu(x)  = \frac{\nu_m}{2 + m} (\wt{\varepsilon}^{2 + m} - \varepsilon^{2 + m}) \int_{\mc{X}} |d_x^1f|^2 \,d\mu(x) = 27\frac{\nu_m}{(2 + m)R^2} \varepsilon^{4 + m} \|d^1f\|_{L^2(\mc{X})}^2.
+\end{equation*}
+Recalling that $\|d^1f\|_{L^2(\mc{X})}^2 \leq \|f\|_{H^1(\mc{X})}^2$, we see that this implies the claim of Lemma~\ref{lem:dirichlet_energy_remainder}.
 
@@
 \label{sec:empirical_norm}
-In this Section we prove Proposition~\ref{prop:empirical_norm_sobolev} (in Section~\ref{subsec:empirical_norm_sobolev}).
-We also prove an analogous result when $\mc{X}$ is a manifold as in Model~\ref{def:model_manifold} (in Section~\ref{subsec:empirical_norm_sobolev_manifold}).
-
-\subsection{Proof of Proposition~\ref{prop:empirical_norm_sobolev}}
+In this Section we prove Proposition~6 (in Section~\ref{subsec:empirical_norm_sobolev}).
+We also prove an analogous result when $\mc{X}$ is a manifold as in Model~2 (in Section~\ref{subsec:empirical_norm_sobolev_manifold}).
+
+\subsection{Proof of Proposition~6}
 \label{subsec:empirical_norm_sobolev}
-In this section we establish Proposition~\ref{prop:empirical_norm_sobolev}.
+In this section we establish Proposition~6.
 As mentioned, the proof of this Proposition follows from the Gagliardo-Nirenberg interpolation inequality, and a one-sided Bernstein's inequality (Lemma~\ref{lem:one_sided_bernstein}).
@@
 	\label{lem:gagliardo_nirenberg}
-	Suppose Model~\ref{def:model_flat_euclidean}, and that $f \in H^s(\mc{X})$ for some $s \geq d/4$. Then there exist constants $C_1$ and $C_2$ that do not depend on $f$, such that
+	Suppose Model~1, and that $f \in H^s(\mc{X})$ for some $s \geq d/4$. Then there exist constants $C_1$ and $C_2$ that do not depend on $f$, such that
 	\begin{equation}
 	\label{eqn:gagliardo_nirenberg}
-	\|f\|_{\Leb^4(\mc{X})} \leq C_1 |f|_{H^s(\mc{X})}^{d/4s} \|f\|_{\Leb^2(\mc{X})}^{1 - d/(4s)} + C_2 \|f\|_{\Leb^2(\mc{X})}
-	\end{equation}
-\end{lemma}
-
-\paragraph{Proof (of Proposition~\ref{prop:empirical_norm_sobolev}).}
+	\|f\|_{L^4(\mc{X})} \leq C_1 |f|_{H^s(\mc{X})}^{d/4s} \|f\|_{L^2(\mc{X})}^{1 - d/(4s)} + C_2 \|f\|_{L^2(\mc{X})}
+	\end{equation}
+\end{lemma}
+
+\paragraph{Proof (of Proposition~6).}
 Rearranging~\eqref{eqn:gagliardo_nirenberg} and raising both sides to the $4$th power, we see that
 \begin{equation*}
-\frac{\Ebb[f^4(X)]}{\|f\|_P^4} \leq C \biggl(\frac{\|f\|_{\Leb^4(\mc{X})}}{\|f\|_{\Leb^2(\mc{X})}}\biggr)^4 \leq C_1\biggl(\frac{|f|_{H^s(\mc{X})}}{\|f\|_{\Leb^2(\mc{X})}}\biggr)^{d/s} + C_2,
+\frac{\Ebb[f^4(X)]}{\|f\|_P^4} \leq C \biggl(\frac{\|f\|_{L^4(\mc{X})}}{\|f\|_{L^2(\mc{X})}}\biggr)^4 \leq C_1\biggl(\frac{|f|_{H^s(\mc{X})}}{\|f\|_{L^2(\mc{X})}}\biggr)^{d/s} + C_2,
 \end{equation*}
 here the constants $C_1,C_2$ are not the same as in~\eqref{eqn:gagliardo_nirenberg}.
-Therefore taking the constant $C$ in assumption~\eqref{eqn:empirical_norm_sobolev_1} to be sufficiently large relative to $C_1$ and $C_2$, we have that
-\begin{equation*}
-C_1\biggl(\frac{|f|_{H^s(\mc{X})}}{\|f\|_{\Leb^2(\mc{X})}}\biggr)^{d/s} \leq \frac{\delta n}{64},
+Therefore taking the constant $C$ in the statement of Proposition~6 to be sufficiently large relative to $C_1$ and $C_2$, we have that
+\begin{equation*}
+C_1\biggl(\frac{|f|_{H^s(\mc{X})}}{\|f\|_{L^2(\mc{X})}}\biggr)^{d/s} \leq \frac{\delta n}{64},
 \end{equation*} 
@@
 \end{equation*}
-The claim then follows from Lemma~\ref{lem:one_sided_bernstein}, upon taking $c = 1/(64C_2^3)$ in the statement of Proposition~\ref{prop:empirical_norm_sobolev}.
-
-\subsection{Proof of Proposition~\ref{prop:empirical_norm_sobolev_manifold}}
+The claim then follows from Lemma~\ref{lem:one_sided_bernstein}, upon taking $c = 1/(64C_2^3)$ in the statement of Proposition~6.
+
+\subsection{Proof of Proposition~9}
 \label{subsec:empirical_norm_sobolev_manifold}
-The proof of Proposition~\ref{prop:empirical_norm_sobolev_manifold} follows exactly the same steps as the proof of Proposition~\ref{prop:empirical_norm_sobolev}, upon replacing Lemma~\ref{lem:gagliardo_nirenberg} by Lemma~\ref{lem:gagliardo_nirenberg_manifold}.
+The proof of Proposition~9 follows exactly the same steps as the proof of Proposition~6, upon replacing Lemma~\ref{lem:gagliardo_nirenberg} by Lemma~\ref{lem:gagliardo_nirenberg_manifold}.
 \begin{lemma}[(c.f Theorem~3.70 of~\citet{aubin2012})]
 	\label{lem:gagliardo_nirenberg_manifold}
-	Suppose Model~\ref{def:model_manifold}, and that $f \in H^s(\mc{X})$ for some $s \geq m/4$. Then there exist constants $C_1$ and $C_2$ that do not depend on $f$, such that
+	Suppose Model~2, and that $f \in H^s(\mc{X})$ for some $s \geq m/4$. Then there exist constants $C_1$ and $C_2$ that do not depend on $f$, such that
 	\begin{equation}
 	\label{eqn:gagliardo_nirenberg_manifold}
-	\|f\|_{\Leb^4(\mc{X})} \leq C_1 |f|_{H^s(\mc{X})}^{m/4s} \|f\|_{\Leb^2(\mc{X})}^{1 - m/(4s)} + C_2 \|f\|_{\Leb^2(\mc{X})}.
+	\|f\|_{L^4(\mc{X})} \leq C_1 |f|_{H^s(\mc{X})}^{m/4s} \|f\|_{L^2(\mc{X})}^{1 - m/(4s)} + C_2 \|f\|_{L^2(\mc{X})}.
 	\end{equation}
@@
 
-\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}.}
+\paragraph{Proof of Theorem~1.}
 We condition on the event that the design points $X_1,\ldots,X_n$ satisfy
@@
 \end{equation}
-Note that by Propositions~\ref{prop:graph_seminorm_fo} and~\ref{prop:graph_eigenvalue}, these statements are both satisfied with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\}$.
+Note that by Propositions~3 and~4, these statements are both satisfied with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\}$.
 
@@
 either deterministically (when $K = 0$), or with probability at least $1 - \exp(-K)$ (when $K \geq 1$).
-Further, from the bounds $\varepsilon \leq c_0 K^{-1/d}$ (Assumption~\ref{asmp:parameters_estimation_fo}) and $\lambda_{K + 1}(\Delta_P) \geq c (K + 1)^{2/d}$ (Weyl's Law) we can simply the above expression to the following,
+Further, from the bounds $\varepsilon \leq c_0 K^{-1/d}$ (Assumption~(P1)) and $\lambda_{K + 1}(\Delta_P) \geq c (K + 1)^{2/d}$ (Weyl's Law) we can simply the above expression to the following,
 \begin{equation}
@@
 \end{equation}
-We now upper bound the right hand side of~\eqref{pf:laplacian_eigenmaps_estimation_fo_2}, based on the value of $K$ chosen in~\ref{asmp:parameters_estimation_fo}.
+We now upper bound the right hand side of~\eqref{pf:laplacian_eigenmaps_estimation_fo_2}, based on the value of $K$ chosen in~(P1).
 When possible we choose $K = \floor{M^2n}^{d/(2 + d)}$ to balance bias and variance, in which case~\eqref{pf:laplacian_eigenmaps_estimation_fo_2} implies
@@
 \end{equation*}
-with probability at least $1 - \exp(-n)$. Combining these three separate cases yields the conclusion of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}.
-
-\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}.}
-Follows identically to the proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~\ref{prop:graph_seminorm_ho} rather than Proposition~\ref{prop:graph_seminorm_fo} and Assumption~\ref{asmp:parameters_estimation_ho} rather than Assumption~\ref{asmp:parameters_estimation_fo}.
-
-\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_manifold}.}
-Follows identically to the proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~\ref{prop:graph_seminorm_manifold} rather than Proposition~\ref{prop:graph_seminorm_fo}, Proposition~\ref{prop:graph_eigenvalue_manifold} rather than Proposition~\ref{prop:graph_eigenvalue}, and Assumption~\ref{asmp:parameters_estimation_manifold} rather than Assumption~\ref{asmp:parameters_testing_fo}.
+with probability at least $1 - \exp(-n)$. Combining these three separate cases yields the conclusion of Theorem~1.
+
+\paragraph{Proof of Theorem~3.}
+Follows identically to the proof of Theorem~1, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~4 rather than Proposition~3 and Assumption~(P3) rather than Assumption~(P1).
+
+\paragraph{Proof of Theorem~6.}
+Follows identically to the proof of Theorem~1, substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~7 rather than Proposition~3, Proposition~8 rather than Proposition~4, and Assumption~(P6) rather than Assumption~(P1).
 
@@
 
-\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}.}
+\paragraph{Proof of Theorem~2.}
 We have already upper bounded the Type I error of $\varphi$ in Lemma~\ref{lem:fixed_graph_testing}, and it remains to upper bound the Type II error.
@@
 \end{equation}
-Note that by Propositions~\ref{prop:graph_seminorm_fo} and~\ref{prop:graph_eigenvalue}, both statements in~\eqref{pf:laplacian_eigenmaps_testing_fo_1} are satisfied with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\}$.
-Additionally, by Proposition~\ref{prop:empirical_norm_sobolev} and the assumption in~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} that $\|f_0\|_P^2 \geq CM^2/(bn^{2/d})$, the one-sided inequality~\eqref{pf:laplacian_eigenmaps_testing_fo_2} follows with probability at least $1 - \exp\{-(cn \wedge 1/b)\}$.
+Note that by Propositions~3 and~4, both statements in~\eqref{pf:laplacian_eigenmaps_testing_fo_1} are satisfied with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\}$.
+Additionally, by Proposition~6 and the assumption that $\|f_0\|_P^2 \geq CM^2/(bn^{2/d})$, the one-sided inequality~\eqref{pf:laplacian_eigenmaps_testing_fo_2} follows with probability at least $1 - \exp\{-(cn \wedge 1/b)\}$.
 Setting $\delta = b/3$ and taking $n \geq N$ to be sufficiently large, the bottom line is that both~\eqref{pf:laplacian_eigenmaps_testing_fo_1} and~\eqref{pf:laplacian_eigenmaps_testing_fo_2} are together satisfied with probability at least $1 - b/2$.
 
-Now, to complete the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}, we would like to invoke Lemma~\ref{lem:fixed_graph_testing}, and conclude that conditional on $X_1,\ldots,X_n$ satisfying~\eqref{pf:laplacian_eigenmaps_testing_fo_1} and~\eqref{pf:laplacian_eigenmaps_testing_fo_2}, our test $\varphi$ will equal $1$ with probability at least $1 - b/2$.
+Now, to complete the proof of Theorem~2, we would like to invoke Lemma~\ref{lem:fixed_graph_testing}, and conclude that conditional on $X_1,\ldots,X_n$ satisfying~\eqref{pf:laplacian_eigenmaps_testing_fo_1} and~\eqref{pf:laplacian_eigenmaps_testing_fo_2}, our test $\varphi$ will equal $1$ with probability at least $1 - b/2$.
 To use Lemma~\ref{lem:fixed_graph_testing}, we will need to establish that~\eqref{eqn:fixed_graph_testing_critical_radius} is satisfied, which we now show.
@@
 On the other hand, we have that $\|f_0\|_n^2 \geq \|f_0\|_P^2/2$.
-Consequently, to prove Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}, it remains only to verify that
+Consequently, to prove Theorem~2, it remains only to verify that
 \begin{equation}
@@
 \end{equation}
-As in the estimation case, we can further upper bound the right hand side of~\eqref{pf:laplacian_eigenmaps_testing_fo_3}, depending on the value of $K$ chosen in~\ref{asmp:parameters_testing_fo}.
+As in the estimation case, we can further upper bound the right hand side of~\eqref{pf:laplacian_eigenmaps_testing_fo_3}, depending on the value of $K$ chosen in~(P2).
 The classical case is $K = (M^2n)^{d/(2 + d)}$, in which case~\eqref{pf:laplacian_eigenmaps_testing_fo_3} is satisfied as long as
@@
 \end{equation*}
-We conclude by observing that~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} implies each of these three inequalities, and thus implies~\eqref{pf:laplacian_eigenmaps_testing_fo_3}.
-
-\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_ho}.}
-Follows identically to the proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~\ref{prop:graph_seminorm_ho} rather than Proposition~\ref{prop:graph_seminorm_fo} and Assumption~\ref{asmp:parameters_testing_ho} rather than Assumption~\ref{asmp:parameters_testing_fo}.
-
-\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_manifold}.}
-Follows identically to the proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~\ref{prop:graph_seminorm_manifold} rather than Proposition~\ref{prop:graph_seminorm_fo}, Proposition~\ref{prop:graph_eigenvalue_manifold} rather than Proposition~\ref{prop:graph_eigenvalue}, Proposition~\ref{prop:empirical_norm_sobolev_manifold} rather than Proposition~\ref{prop:empirical_norm_sobolev}, and Assumption~\ref{asmp:parameters_testing_manifold} rather than Assumption~\ref{asmp:parameters_testing_fo}.
-
-\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}.}
+We conclude by observing that by assumption, $\|f_0\|_P^2$ implies each of these three inequalities, and thus implies~\eqref{pf:laplacian_eigenmaps_testing_fo_3}.
+
+\paragraph{Proof of Theorem~4.}
+Follows identically to the proof of Theorem~1, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~4 rather than Proposition~3 and Assumption~(P4) rather than Assumption~(P2).
+
+\paragraph{Proof of Theorem~7.}
+Follows identically to the proof of Theorem~1, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~7 rather than Proposition~3, Proposition~8 rather than Proposition~4, Proposition~9 rather than Proposition~6, and Assumption~(P6) rather than Assumption~(P2).
+
+\paragraph{Proof of Theorem~5.}
 Note that our choices of $K$ and $\varepsilon$ ensure  that~\eqref{pf:laplacian_eigenmaps_testing_fo_1} (with $L_{n,\varepsilon}^s$ replacing $L_{n,\varepsilon}$) and~\eqref{pf:laplacian_eigenmaps_testing_fo_2} are satisfied with probability at least $1 - b/2$.
-Proceeding as in the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}, we upper bound the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius},
+Proceeding as in the proof of Theorem~2, we upper bound the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius},
 \begin{align*}
@@
 \end{align*}
-Unlike in the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}, we note that in this case $\varepsilon^2 \leq C\lambda_K(\Delta_P)$ rather than vice versa.
-From here, proceeding as in the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} gives the claimed result.
+Unlike in the proof of Theorem~2, we note that in this case $\varepsilon^2 \leq C\lambda_K(\Delta_P)$ rather than vice versa.
+From here, proceeding as in the proof of Theorem~2 gives the claimed result.
 
@@
 \label{subsec:kernel_smoothing}
-In this section we prove Lemma~\ref{lem:kernel_smoothing_insample} (in Section~\ref{subsec:pf_kernel_smoothing_insample}) and Lemma~\ref{lem:kernel_smoothing_bias} (in Section~\ref{subsec:pf_kernel_smoothing_bias}).
+In this section we prove Lemma~1 (in Section~\ref{subsec:pf_kernel_smoothing_insample}) and Lemma~\ref{lem:kernel_smoothing_bias} (in Section~\ref{subsec:pf_kernel_smoothing_bias}).
 In Section~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we give a sequence of design densities and regression functions $\{p^{(n)}(x), f_0^{(n)}(x)\}_{n \in \mathbb{N}}$ for which the estimator $T_{n,h}\wc{f}$ (extension of Laplacian eigenmaps by kernel smoothing) strictly outperforms directly kernel smoothing the responses, in the sense that
@@
 This expression reflects the known fact that the bias operator of kernel smoothing is equal to the non-local Laplacian, up to a rescaling by the population degree functional $d_{P,h}(x)$.
-In the second equality, we are using the notation $L_{P,h}f(x)$ exactly as defined in~\eqref{eqn:nonlocal_laplacian}, but with the kernel $\psi$ instead of $\eta$.
+In the second equality, we are using the notation $L_{P,h}f(x)$ exactly as defined in equation~(35) of the main text, but with the kernel $\psi$ instead of $\eta$.
 Note that $\psi$ satisfies all the same assumptions as $\eta$, except that of positivity; when $\psi$ is a higher-order kernel it may take negative values.
@@
 \end{align*}
-Since by assumption $\mc{X}$ has Lipschitz boundary, setting $c_0$ to be a sufficiently small constant in~\ref{asmp:bandwidth}, we can further deduce that $\int \psi(\|z\|) \1\{hz + x \in \mc{X}\} \,dz \geq 1/3$, and consequently that
+Since by assumption $\mc{X}$ has Lipschitz boundary, setting $c_0$ to be a sufficiently small constant in~(P7), we can further deduce that $\int \psi(\|z\|) \1\{hz + x \in \mc{X}\} \,dz \geq 1/3$, and consequently that
 \begin{equation}
@@
 
-\subsection{Proof of Lemma~\ref{lem:kernel_smoothing_insample}}
+\subsection{Proof of Lemma~1}
 \label{subsec:pf_kernel_smoothing_insample}
@@
 	\label{lem:kernel_smoothing_bias}
-	Suppose Model~\ref{def:model_flat_euclidean}, and that the kernel smoothing operator $T_{P,h}$ is computed with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel}.
+	Suppose Model~1, and that the kernel smoothing operator $T_{P,h}$ is computed with respect to a kernel $\eta$ that satisfies~(K3).
 	\begin{itemize}
@@
 		\end{equation*}
-		\item If $f_0 \in H_0^{s}(\mc{X})$, $p \in C^{s - 1}(\mc{X})$, and $\eta$ satisfies~\ref{asmp:ho_kernel}, then there exists a constant $C$ which does not depend on $f_0$ such that
+		\item If $f_0 \in H_0^{s}(\mc{X})$, $p \in C^{s - 1}(\mc{X})$, and $\eta$ satisfies~(K4), then there exists a constant $C$ which does not depend on $f_0$ such that
 		\begin{equation*}
@@
 \begin{align*}
-\|T_{n,h}\wh{f} - f_0\|_{\Leb^2(P^{(n)})}^2 & \leq 2 \|f_0\|_{\Leb^2(P^{(n)})}^2 + \sum_{i = 1}^{n} (\wh{f}_i)^2 \biggl(\int \frac{\1\{\|X_i - x\| \leq h \}}{d_{n,h}(x)} \,dP(x)\biggr) \\
-& \leq 2 \|f_0\|_{\Leb^2(P^{(n)})}^2 + 2h\sum_{i = 1}^{n} (\wh{f}_i)^2 \\
-& = 2 \|f_0\|_{\Leb^2(P^{(n)})}^2 + 2 (\log n)^2 \|\wh{f}\|_n^2.
+\|T_{n,h}\wh{f} - f_0\|_{L^2(P^{(n)})}^2 & \leq 2 \|f_0\|_{L^2(P^{(n)})}^2 + \sum_{i = 1}^{n} (\wh{f}_i)^2 \biggl(\int \frac{\1\{\|X_i - x\| \leq h \}}{d_{n,h}(x)} \,dP(x)\biggr) \\
+& \leq 2 \|f_0\|_{L^2(P^{(n)})}^2 + 2h\sum_{i = 1}^{n} (\wh{f}_i)^2 \\
+& = 2 \|f_0\|_{L^2(P^{(n)})}^2 + 2 (\log n)^2 \|\wh{f}\|_n^2.
 \end{align*}
@@
 \begin{equation*}
-\Ebb\Bigl[\|T_{n,h}\wh{f} - f_0\|_{\Leb^2(P^{(n)})}^2 \cdot \1\{\mc{E}^c\}\Bigr] \leq \Ebb\Bigl[\Bigl(2 \|f_0\|_{\Leb^2(P^{(n)})}^2 + 4 (\log n)^2 \|{\bf Y}\|_n^2\Bigr) \cdot \1\{\mc{E}^c\} \Bigr] \leq \frac{2 \theta^2}{n} + \frac{4 (\log n)^2(\theta^2 + 1)}{n}.
+\Ebb\Bigl[\|T_{n,h}\wh{f} - f_0\|_{L^2(P^{(n)})}^2 \cdot \1\{\mc{E}^c\}\Bigr] \leq \Ebb\Bigl[\Bigl(2 \|f_0\|_{L^2(P^{(n)})}^2 + 4 (\log n)^2 \|{\bf Y}\|_n^2\Bigr) \cdot \1\{\mc{E}^c\} \Bigr] \leq \frac{2 \theta^2}{n} + \frac{4 (\log n)^2(\theta^2 + 1)}{n}.
 \end{equation*}
@@
 	\begin{equation*}
-	\Pbb\Bigl[U \geq 2 \norm{a}_2 \sqrt{t} + 2 \norm{a}_{\infty}t\Bigr] \leq \exp(-t).
+	\Pbb\Bigl[U \geq 2 \|a\|_2 \sqrt{t} + 2 \|a\|_{\infty}t\Bigr] \leq \exp(-t).
 	\end{equation*}
@@
 
-In the proof of Lemma~\ref{lem:kernel_smoothing_insample}, we require uniform control of the empirical degree functional $d_{n,h}(x)$ over all $x \in \mc{X}$.
+In the proof of Lemma~1, we require uniform control of the empirical degree functional $d_{n,h}(x)$ over all $x \in \mc{X}$.
 Such a result is available to us because the kernel $\psi$ is Lipschitz on its support, so that the class of functions $\{\psi((x - \cdot)/h): h \in \Rd\}$ has finite VC dimension.
@@
 	\label{lem:uniform_bound_empirical_degree}
-	Suppose Model~\ref{def:model_flat_euclidean}.
-For a kernel $\psi$ satisfying~\ref{asmp:kernel} and bandwidth $h$ satisfying~\ref{asmp:bandwidth}, there exist constants $c, C, c_1$ and $C_1$ which do not depend on $h$ or $n$ such that
+	Suppose Model~1. For a kernel $\psi$ satisfying~(K3) and bandwidth $h$ satisfying~(P7), there exist constants $c, C, c_1$ and $C_1$ which do not depend on $h$ or $n$ such that
 	\begin{equation*}
@@
 Recall the lower bound on $d_{P,h}(x)$ given in~\eqref{eqn:degree_lower_bound}.
-By setting $C_0$ to be a sufficiently large constant in~\ref{asmp:bandwidth}, we can ensure that choosing $t = n h^d p_{\min}/6$ satisfies both the inequalities~\eqref{eqn:uniform_bound_empirical_degree_1}.
+By setting $C_0$ to be a sufficiently large constant in~(P7), we can ensure that choosing $t = n h^d p_{\min}/6$ satisfies both the inequalities~\eqref{eqn:uniform_bound_empirical_degree_1}.
 For this choice of $t$ it follows from~\eqref{eqn:degree_lower_bound} and Lemma~\ref{lem:uniform_bound_empirical_degree} that
@@
 \end{equation}
-with probability at least $1 - C\exp(-cnh^d)$. This is the form of the result we use in the proof of Lemma~\ref{lem:kernel_smoothing_insample}.
+with probability at least $1 - C\exp(-cnh^d)$. This is the form of the result we use in the proof of Lemma~1.
 
@@
 We begin with some notation that allows us to concisely derivatives.
-For a given $z \in \Rd$ and $s$-times differentiable function $f: \mc{X} \to \Reals$, we denote $\bigl(d_x^sf\bigr)(z) := \sum_{\abs{\alpha} = s} D^{\alpha}f(x) z^{\alpha}$.
-We also write $d^sf := \sum_{\abs{\alpha} = j} D^{\alpha}f$.
+For a given $z \in \Rd$ and $s$-times differentiable function $f: \mc{X} \to \Reals$, we denote $\bigl(d_x^sf\bigr)(z) := \sum_{|\alpha| = s} D^{\alpha}f(x) z^{\alpha}$.
+We also write $d^sf := \sum_{|\alpha| = j} D^{\alpha}f$.
 We point out that in the first-order case $d_x^1f$ is the differential of $f$ at $x \in \mc{X}$, while $d^1f$ is the divergence of $f$.
@@
 \label{eqn:sobolev_remainder_term}
-\int_{\mc{X}_{h}} \bigl|r_{x + thz}^j(x;f)\bigr|^2 \,dx \leq h^{2j} \int_{\mc{X}_{h}} \int_{0}^{1} |d_{x + thz}^jf(z)|^2 \,dt \,dx \leq h^{2j} \|d^jf\|_{\Leb^2(\mc{X})}^2.
+\int_{\mc{X}_{h}} \bigl|r_{x + thz}^j(x;f)\bigr|^2 \,dx \leq h^{2j} \int_{\mc{X}_{h}} \int_{0}^{1} |d_{x + thz}^jf(z)|^2 \,dt \,dx \leq h^{2j} \|d^jf\|_{L^2(\mc{X})}^2.
 \end{equation}
@@
 As pointed out by~\citet{sadhanala16b}, in this special case there are far simpler and faster methods for sparsification, which at least empirically seem to do the job.
+
+% Bibliography
+\bibliographystyle{imsart-nameyear}
+\bibliography{../../../graph_regression_bibliography}
+
+\end{document}

### 2021-09-04 manuscripts/laplacian_eigenmaps/arxiv/laplacian_eigenmaps_regression.tex
Put Laplacian Eigenmaps paper into AOS format.

--- 2021-08-29 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
@@
 \documentclass{article}
-\usepackage{amsmath}
-\usepackage{amsfonts, amsthm, amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\hypersetup{
-	colorlinks=true,
-	linkcolor=blue,
-	citecolor=blue
+
+%%% Begin Ryan's template
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{times}          % times font
+
+\usepackage[round]{natbib}
+\usepackage{amssymb,amsmath,amsthm,bbm}
+\usepackage[margin=1in]{geometry}
+\usepackage{verbatim,float,url,dsfont}
+\usepackage{graphicx,subfigure,psfrag}
+\usepackage{algorithm,algorithmic}
+\usepackage{mathtools,enumitem}
+\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}
+\usepackage{multirow}
+
+% Theorems and such
+\newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
+\newtheorem{corollary}{Corollary}
+\newtheorem{proposition}{Proposition}
+\theoremstyle{definition}
+\newtheorem{remark}{Remark}
+\newtheorem{definition}{Definition}
+
+% Assumption
+\newtheorem*{assumption*}{\assumptionnumber}
+\providecommand{\assumptionnumber}{}
+\makeatletter
+\newenvironment{assumption}[2]{
+	\renewcommand{\assumptionnumber}{Assumption #1#2}
+	\begin{assumption*}
+		\protected@edef\@currentlabel{#1#2}}
+	{\end{assumption*}}
+\makeatother
+
+% Widebar
+\makeatletter
+\newcommand*\rel@kern[1]{\kern#1\dimexpr\macc@kerna}
+\newcommand*\widebar[1]{%
+	\begingroup
+	\def\mathaccent##1##2{%
+		\rel@kern{0.8}%
+		\overline{\rel@kern{-0.8}\macc@nucleus\rel@kern{0.2}}%
+		\rel@kern{-0.2}%
+	}%
+	\macc@depth\@ne
+	\let\math@bgroup\@empty \let\math@egroup\macc@set@skewchar
+	\mathsurround\z@ \frozen@everymath{\mathgroup\macc@group\relax}%
+	\macc@set@skewchar\relax
+	\let\mathaccentV\macc@nested@a
+	\macc@nested@a\relax111{#1}%
+	\endgroup
 }
-\usepackage[parfill]{parskip}
-\usepackage{algpseudocode}
-\usepackage{algorithm}
-\usepackage{enumerate}
-\usepackage[shortlabels]{enumitem}
-\usepackage{fullpage}
-\usepackage{mathtools}
-\usepackage{tikz}
-\usepackage{bm}
-\usepackage{multirow}
-\usepackage[font={small,it}]{caption}
+\makeatother
 
-\usepackage{natbib}
-\renewcommand{\bibname}{References}
-\renewcommand{\bibsection}{\subsubsection*{\bibname}}
+% Min and max
+\newcommand{\argmin}{\mathop{\mathrm{argmin}}}
+\newcommand{\argmax}{\mathop{\mathrm{argmax}}}
+\newcommand{\minimize}{\mathop{\mathrm{minimize}}}
+\newcommand{\st}{\mathop{\mathrm{subject\,\,to}}}
+
+% Shortcuts
+\def\R{\mathbb{R}}
+
+%%% End Ryan's template
+
+%%% Begin Alden's additions
+\newcommand{\Ebb}{\mathbb{E}}
+\newcommand{\Pbb}{\mathbb{P}}
+\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
+\newcommand{\wt}[1]{\widetilde{#1}}
+\newcommand{\wh}[1]{\widehat{#1}}
+\newcommand{\mc}[1]{\mathcal{#1}}
+\newcommand{\Reals}{\mathbb{R}} % Same thing as Ryan's \R
+\newcommand{\Rd}{\Reals^d}
+\newcommand{\wb}[1]{\widebar{#1}}
+\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
+\newcommand{\Var}{\mathrm{Var}}
+\newcommand{\Cov}{\mathrm{Cov}}
+\newcommand{\1}{\mathbf{1}}
+\newcommand{\bj}{{\bf j}}
+\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
 
@@
 \DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
-\DeclareMathAccent{\wb}{0}{mathx}{"73}
 \DeclareMathAccent{\wc}{0}{mathx}{"71}
-
-\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
-\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}
-
-\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
-\newcommand{\Graph}{\mathcal{G}}
-\newcommand{\Reals}{\mathbb{R}}
-\newcommand{\Identity}{\mathbb{I}}
-\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
-\newcommand{\convprob}{\overset{p}{\to}}
-\newcommand{\convdist}{\overset{w}{\to}}
-\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
-\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
-\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
-\newcommand{\iset}{\mathbf{i}}
-\newcommand{\jset}{\mathbf{j}}
-\newcommand{\myexp}[1]{\exp \{ #1 \}}
-\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
-\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
-\newcommand{\ext}[1]{\widetilde{#1}}
-\newcommand{\set}[1]{\left\{#1\right\}}
-\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
-\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
-\newcommand{\Var}{\mathrm{Var}}
-\newcommand{\Cov}{\mathrm{Cov}}
-\newcommand{\diam}{\mathrm{diam}}
-
-\newcommand{\emC}{C_n}
-\newcommand{\emCpr}{C'_n}
-\newcommand{\emCthick}{C^{\sigma}_n}
-\newcommand{\emCprthick}{C'^{\sigma}_n}
-\newcommand{\emS}{S^{\sigma}_n}
-\newcommand{\estC}{\widehat{C}_n}
-\newcommand{\hC}{\hat{C^{\sigma}_n}}
-\newcommand{\vol}{\text{vol}}
-\newcommand{\spansp}{\mathrm{span}~}
-\newcommand{\1}{\mathbf{1}}
-
-\newcommand{\Linv}{L^{\Xsetagger}}
-\DeclareMathOperator*{\argmin}{argmin}
-\DeclareMathOperator*{\argmax}{argmax}
-
-\newcommand{\emF}{\mathbb{F}_n}
-\newcommand{\emG}{\mathbb{G}_n}
-\newcommand{\emP}{\mathbb{P}_n}
-\newcommand{\F}{\mathcal{F}}
-\newcommand{\D}{\mathcal{D}}
-\newcommand{\R}{\mathcal{R}}
-\newcommand{\Rd}{\Reals^d}
-\newcommand{\RD}{\Reals^D}
-\newcommand{\Nbb}{\mathbb{N}}
-
-%%% Vectors
-\newcommand{\thetast}{\theta^{\star}}
-\newcommand{\betap}{\beta^{(p)}}
-\newcommand{\betaq}{\beta^{(q)}}
-\newcommand{\vardeltapq}{\varDelta^{(p,q)}}
-\newcommand{\lambdavec}{\boldsymbol{\lambda}}
-\newcommand{\bj}{{\bf j}}
-
-%%% Matrices
-\newcommand{\X}{X} % no bold
-\newcommand{\Y}{Y} % no bold
-\newcommand{\Z}{Z} % no bold
-\newcommand{\Lgrid}{L_{\grid}}
-\newcommand{\Xsetgrid}{D_{\grid}}
-\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}
-\newcommand{\Lap}{{\bf L}}
-\newcommand{\NLap}{{\bf N}}
-\newcommand{\PLap}{{\bf P}}
-\newcommand{\Id}{I}
-
-%%% Sets and classes
-\newcommand{\Xset}{\mathcal{X}}
-\newcommand{\Sset}{\mathcal{S}}
-\newcommand{\Hclass}{\mathcal{H}}
-\newcommand{\Pclass}{\mathcal{P}}
-\newcommand{\Leb}{L}
-\newcommand{\mc}[1]{\mathcal{#1}}
-
-%%% Distributions and related quantities
-\newcommand{\Pbb}{\mathbb{P}}
-\newcommand{\Ebb}{\mathbb{E}}
-\newcommand{\Qbb}{\mathbb{Q}}
-\newcommand{\Ibb}{\mathbb{I}}
-
-%%% Operators
-\newcommand{\Tadj}{T^{\star}}
-\newcommand{\Xsetive}{\mathrm{div}}
-\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
-\newcommand{\gradient}{\mathcal{D}}
-\newcommand{\Hessian}{\mathcal{D}^2}
-\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
-\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}
-
-%%% Misc
-\newcommand{\grid}{\mathrm{grid}}
-\newcommand{\critr}{R_n}
-\newcommand{\Xsetx}{\,dx}
-\newcommand{\Xsety}{\,dy}
-\newcommand{\Xsetr}{\,dr}
-\newcommand{\Xsetxpr}{\,dx'}
-\newcommand{\Xsetypr}{\,dy'}
-\newcommand{\wt}[1]{\widetilde{#1}}
-\newcommand{\wh}[1]{\widehat{#1}}
-\newcommand{\ol}[1]{\overline{#1}}
-\newcommand{\spec}{\mathrm{spec}}
-\newcommand{\LE}{\mathrm{LE}}
-\newcommand{\LS}{\mathrm{LS}}
-\newcommand{\SM}{\mathrm{SM}}
-\newcommand{\OS}{\mathrm{FS}}
-\newcommand{\PLS}{\mathrm{PLS}}
-
-%%% Order of magnitude
-\newcommand{\soom}{\sim}
-
-% \newcommand{\span}{\textrm{span}}
-
-\newtheoremstyle{alden}
-{6pt} % Space above
-{6pt} % Space below
-{} % Body font
-{} % Indent amount
-{\bfseries} % Theorem head font
-{.} % Punctuation after theorem head
-{.5em} % Space after theorem head
-{} % Theorem head spec (can be left empty, meaning `normal')
-
-\theoremstyle{alden}
-
-\newtheoremstyle{aldenthm}
-{6pt} % Space above
-{6pt} % Space below
-{\itshape} % Body font
-{} % Indent amount
-{\bfseries} % Theorem head font
-{.} % Punctuation after theorem head
-{.5em} % Space after theorem head
-{} % Theorem head spec (can be left empty, meaning `normal')
-
-\theoremstyle{aldenthm}
-\newtheorem{theorem}{Theorem}
-\newtheorem{conjecture}{Conjecture}
-\newtheorem{lemma}{Lemma}
-\newtheorem{example}{Example}
-\newtheorem{corollary}{Corollary}
-\newtheorem{proposition}{Proposition}
-\newtheorem{assumption}{Assumption}
-\newtheorem{remark}{Remark}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-
-\theoremstyle{remark}
+%%% End Alden's Additions
 
 \begin{document}
-\title{Minimax-optimal Laplacian Eigenmaps regression over Sobolev Spaces with Neighborhood Graphs}
-\author{Alden Green}
-\date{\today}
-\maketitle
 
-\input{sections/introduction}
-\input{sections/preliminaries}
+% Title 
+\begin{center} {\Large{\bf{Minimax-optimal Regression over Sobolev Spaces \\
+				\vspace{.2cm}
+				via Laplacian Eigenmaps with Neighborhood Graphs}}}
 
-\section{Minimax Optimality of Laplacian Eigenmaps}
-\label{sec:minimax_optimal_laplacian_eigenmaps}
+	\vspace*{.3cm}
 
-% As previously explained, Laplacian eigenmaps is a discrete and noisy approximation to a spectral projection method using the eigenfunctions of $\Delta_P$.
-This is particularly useful when $P$ is unknown, or when the eigenfunctions of $\Delta_P$ cannot be explicitly computed.
-Our goal is to show that Laplacian eigenmaps methods are rate-optimal, notwithstanding the potential extra error incurred by this approximation.
-In this section and the following one, we will see that this is indeed the case: the estimator $\wh{f}$, and a test using the statistic $\wh{T}$, achieve optimal estimation and goodness-of-fit testing rates over Sobolev classes.
-% (AG 8/26/21): This segment is now redundant.
+	{\large{
+			\begin{center}
+				Alden Green~~~~~ Sivaraman Balakrishnan~~~~~ Ryan J. Tibshirani\\
+				\vspace{.2cm}
+			\end{center}
 
-In this section we give upper bounds on the error of Laplacian Eigenmaps in the flat Euclidean case, where we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to Model~\ref{def:model_flat_euclidean}.
-We will divide our theorem statements based on the regression function $f_0$ belongs to the first order Sobolev class $H^1(\mc{X})$ or a higher-order Sobolev class ($H_0^{s}(\mc{X})$ for some $s > 1$), since the details of the two settings are somewhat different.
+			\begin{tabular}{c}
+				Department of Statistics and Data Science \\
+				Carnegie Mellon University
+			\end{tabular}
 
-\subsection{First-order Sobolev classes}
-\label{sec:first_order_sobolev_classes}
-We begin by assuming $f_0 \in H^1(\mc{X}; M)$.
-We show that $\wh{f}$ and a test based on $\wh{T}$ are minimax optimal, for all values of $d$ for which the minimax rates are known, and under no additional assumptions (beyond those of Model~\ref{def:model_flat_euclidean}) on the design distribution $P$.
+			\vspace*{.2in}
 
-\paragraph{Estimation.} Laplacian eigenmaps depends on the kernel $\eta$ and two tuning parameters, the graph radius $\varepsilon$ and number of eigenvectors $K$. We now give some assumptions on each.
-\begin{enumerate}[label=(K\arabic*)]
-	\setcounter{enumi}{0}
-	\item
-	\label{asmp:kernel_flat_euclidean}
-	The kernel function $\eta$ is a nonincreasing function supported on $[0,1]$. Its restriction to $[0,1]$ is Lipschitz, and $\eta(1) > 0$. Additionally, it is normalized so that
-	\begin{equation*}
-	\int_{\Rd} \eta(\|z\|) \,dz = 1.
-	\end{equation*}
-	and we assume \smash{$\sigma_{\eta} := \frac{1}{d}\int_{\Rd} \|x\|^2 \eta(\|x\|) \,dx < \infty$}.
-\end{enumerate}
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{0}
-	\item 
-	\label{asmp:parameters_estimation_fo} 
-	For constants $c_0$ and $C_0$, the graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy the following inequalities:
-	\begin{equation}\\
-	\label{eqn:radius_fo} 
-	C_0\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq \varepsilon \leq c_0\min\{1,K^{-1/d}\},
-	\end{equation}
-	and 
-	\begin{equation}
-	\label{eqn:eigenvector_estimation_fo} 
-	K = \min\Bigl\{\floor{(M^2n)^{d/(2 + d)}} \vee 1, n\Bigr\}.
-	\end{equation}
-\end{enumerate}
-We comment on these assumptions after stating our first main theorem, regarding the estimation error of Laplacian eigenmaps.
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_estimation_fo}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H^1(\mc{X},M)$.
-There are constants $c,C$ and $N$ (not depending on $f_0$, $M$ or $n$), such that the following statement holds for all $n \geq N$ and any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_fo}, then
-	\begin{equation}
-	\label{eqn:laplacian_eigenmaps_estimation_fo}
-	\|\wh{f} - f_0\|_n^2 \leq C\Bigl(\frac{1}{\delta}M^2(M^2n)^{-2/(2 + d)} \wedge 1\Bigr) \vee \frac{1}{n},
-	\end{equation}
-	with probability at least $1 - \delta - Cn\exp(-cn\varepsilon^d) - \exp(-K)$.
-\end{theorem}
-From~\eqref{eqn:laplacian_eigenmaps_estimation_fo} it follows immediately that when $M \asymp 1$, then with constant probability $\|\wh{f} - f_0\|_n^2 \lesssim n^{-2/(2 + d)}$, matching the standard minimax estimation rate over Sobolev classes.
+			\begin{tabular}{c}
+				\texttt{\{ajgreen,siva,ryantibs\}@stat.cmu.edu}
+			\end{tabular}
+	}}
 
-Some other remarks:
-\begin{itemize}
-	\item \emph{Radius of the Sobolev ball.} If $M = M_n$ satisfies $n^{-1/2} \lesssim M_n \lesssim n^{1/d}$, then Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} implies that $\|\wh{f} - f_0\|_n^2 \lesssim M^2(M^2n)^{-2/(2 + d)}$ with constant probability.
-This matches the upper bound on $\Ebb\|\wt{f} - f_0\|_P^2$ given in~\eqref{eqn:spectral_series_estimation}, including in its dependence on $M$.
-It also matches the known minimax rate for some related problems, including estimation in the Gaussian sequence model~\citep{johnstone2011}, and estimation with a fixed design when the dimension $d = 1$~\citep{vandergeer2000}.
+	\vspace*{.2in}
 
-	When $M = o(n^{-1/2})$ then computing Laplacian eigenmaps with $K = 1$ achieves the parametric rate $\|\wh{f} - f_0\|_n^2 \lesssim n^{-1}$, and the zero-estimator $\wh{f} = 0$ achieves the better rate $\|\wh{f} - f_0\|_n^2 \lesssim M^2$.
-However, we do not know what the minimax rate is in this regime.
-On the other hand, when $M = \omega(n^{1/d})$, then computing Laplacian eigenmaps with $K = n$ achieves the rate $\|\wh{f} - f_0\|_n^2 \lesssim 1$, which is better than the rate in~\eqref{eqn:spectral_series_estimation}.
-This is because we are evaluating error in-sample rather than out-of-sample.
-However, in truth these are edge cases, which do not fall neatly into the framework of nonparametric regression.
+	\today
+	\vspace*{.2in}
+\end{center}
 
-	\item \emph{In-sample error.} As we've already noted, since the Laplacian eigenmaps estimator is defined only at the design points $\{X_1,\ldots,X_n\}$, we will use the empirical norm $\|\cdot\|_n^2$ as our estimation loss, and return to the question of out-of-sample estimation later in Section~\ref{sec:out_of_sample}.
+% Abstract
+\begin{abstract}
+	In this paper we study the statistical properties of nonparametric regression methods based on \emph{Laplacian eigenmaps}.
+These are spectral projection methods, which involve projecting a vector of observed responses ${\bf Y} = (Y_1,\ldots,Y_n)$ onto a subspace spanned by certain eigenvectors of a neighborhood graph Laplacian.
+We show that in many cases these methods achieve minimax rates of convergence over continuum function classes.
+Specifically, we consider the problem of random design regression over Sobolev spaces $H^s(\mc{X})$---where $\mc{X} \subseteq \Rd$ is the support of the design density $p$---and show that under sufficient smoothness conditions on $p$, methods based on Laplacian eigenmaps achieve the optimal rates for both estimation (where the optimal rate is known to be $n^{-2s/(2s + d)}$) and goodness-of-fit testing ($n^{-4s/(4s + d)}$).
+We also show that Laplacian eigenmaps is \emph{manifold adaptive}: that is, we consider the situation where $\mc{X}$ is a manifold of small intrinsic dimension $m$, and give upper bounds establishing that Laplacian eigenmaps achieves the faster minimax estimation ($n^{-2s/(2s + m)}$) and testing ($n^{-4s/(4s + m)}$) rates of convergence.
+Finally, since the Laplacian eigenmaps estimator is defined only at observed design points, we propose an out-of-sample extension based on kernel smoothing, and prove that it achieves optimal out-of-sample error.
+We support these theoretical results with empirical evidence.
+\end{abstract}
 
-	There is one subtlety introduced by the use of in-sample mean squared error.
-Since elements $f \in H^s(\mc{X})$ are equivalence classes, defined only up to a set of measure zero, one cannot really speak of the pointwise evaluation $f_0(X_i)$, as we do by defining our target of estimation to be $(f_0(X_1),\ldots,f_0(X_n))$, until one selects a representative of each equivalence class $f$.
-Implicitly, we will always pick the \emph{precise representative} $f_0^{\ast} \in f_0$ (as defined in~\cite{evans15}), and the notation ``$f_0(X_i)$'' should always be interpreted as $f_0^{\ast}(X_i)$.
-To be clear, however, it does not really matter which representative we choose, since all versions agree except on a set of measure zero, and so any two $g_0,h_0 \in f_0$ satisfy $g_0(X_i) = h_0(X_i)$ for all $i = 1,\ldots,n$ almost surely.
-For this reason we can write $f_0(X_i)$ without fear of ambiguity or confusion.
+% Main text
+\input{../main_text.tex}
 
-	\item \emph{Tuning parameters}.
-The assumptions placed on the kernel function $\eta$ are needed for technical reasons.
-They can likely be weakened, although we note that they are already fairly general.
-The lower bound on $\varepsilon$ imposed by~\eqref{eqn:radius_fo} is on the order of the connectivity threshold, the smallest radius for which the resulting graph will still be connected with high probability.
-On the other hand, as we will see in Section~\ref{subsec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K$ is of at least the same order as the continuum eigenvalue $\lambda_K(\Delta_P)$; this is essential in order to obtain a tight upper bound on the bias of $\wh{f}$.
-Finally, we set $K = \floor{(M^2n)^{d/(2 + d)}}$ (when possible) to optimally trade-off bias and variance.
+% Bibliography
+\bibliographystyle{plainnat}
+\bibliography{../../../graph_regression_bibliography}
 
-	In practice, one typically tunes hyper-parameters by cross-validation.
-However, because the estimator $\wh{f}$ is defined only in-sample, neither cross-validation nor any other sample-splitting method can be used to tune parameters for Laplacian eigenmaps.
-We return to this issue in Section~\ref{sec:out_of_sample}, when we propose an out-of-sample extension of $\wh{f}$.
-
-	% (AG 8/27/21): Ryan, you asked about AIC, but I'm not clear about whether you want me to (a) make some remark about it, (b) actually go through the work of giving guarantees, or (c) were just curious.
-	\item \emph{High-probability guarantees}.
-The upper bound given in~\eqref{eqn:laplacian_eigenmaps_estimation_fo} holds with ``constant probability'', meaning with probability $1 - \delta - o(1)$.
-Under the stronger assumption that $f_0$ is $M$-Lipschitz, we can establish the same guarantee~\eqref{eqn:laplacian_eigenmaps_estimation_fo} with probability $1 - \delta^2/n - Cn\exp(-cn\varepsilon^d) - \exp(-K)$; in other words, we can give a high probability guarantee (for details see~\citep{green2021}).
-In this case a routine calculation shows that $\Ebb[\|\wh{f} - f_0\|_n^2]$ will also be on the some order as~\eqref{eqn:laplacian_eigenmaps_estimation_fo}.
-We also suspect that high-probability guarantees will hold so long as $\|\nabla f\|_{L^q(\mc{X})}$ is bounded for some sufficiently large $q < \infty$, but it remains an open question whether such guarantees can be obtained in the Sobolev case ($q = 2$) which is the focus of this work.
-\end{itemize}
-
-\paragraph{Testing.} Consider the test $\varphi = \1\{\wh{T} \geq t_{a}\}$, where $t_{a}$ is the threshold
-\begin{equation*}
-t_{a} := \frac{K}{n} + \frac{1}{n}\sqrt{\frac{2K}{a}}.
-\end{equation*}
-This choice of threshold $t_{a}$ guarantees that $\varphi$ is a level-$a$ test.
-As we show in Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}, when $d < 4$, $\varepsilon$ and $K$ are chosen appropriately, and the alternative $f_0$ has is sufficiently well-separated from $0$, the test $\varphi$ has Type II error of at most $b$.
-
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{1}
-	\item 
-	\label{asmp:parameters_testing_fo}
-	The graph radius $\varepsilon$ satisfies~\eqref{eqn:radius_fo}, and the number of eigenvectors 
-	\begin{equation}
-	\label{eqn:eigenvector_testing_fo}
-	K = \min\Bigl\{\floor{(M^2n)^{2d/(4 + d)}} \vee 1, n\Bigr\}.
-	\end{equation}
-\end{enumerate}
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_testing_fo}
-	Fix $a,b \in (0,1)$.
-Suppose Model~\ref{def:model_flat_euclidean}.
-Then $\mathbb{E}_0[\varphi] \leq a$, i.e $\varphi$ is a level-$a$ test.
-Suppose additionally $f_0 \in H^1(\mc{X},M)$, and that $d < 4$.
-Then there exist constants $C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_fo}, and if $f_0$ satisfies
-	\begin{equation}
-	\label{eqn:laplacian_eigenmaps_testing_criticalradius_fo}
-	\|f_0\|_P^2 \geq C\biggl(\Bigl(M^2(M^2n)^{-4/(4 + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{2/d}} \biggr) \vee \frac{1}{n},
-	\end{equation}
-	then $\Ebb_{f_0}[1 - \phi] \leq b$.
-\end{theorem}
-Although~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} involves taking the maximum of several different terms, the important takeaway of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} is that if $n^{-1/2} \lesssim M \lesssim n^{(4 - d)/4d}$, then $\varphi$ has small worst-case risk as long as $f_0$ is separated from $0$ by at least $M^2(M^2n)^{-4/(4 + d)}$.
-In the special case $M \asymp 1$, this matches the minimax squared critical radius $n^{-4/(4 + d)}$, implying that $\varphi$ is a minimax rate-optimal test over $H^1(\mc{X};M)$ when $d \in \{1,2,3\}$.
-As mentioned previously, when $d \geq 4$ the first order Sobolev space $H^1(\mc{X})$ does not continuously embed into $\Leb^4(\mc{X})$, and in this case the optimal rates for regression testing over Sobolev spaces are unknown.
-
-\subsection{Higher-order Sobolev classes}
-\label{sec:higher_order_sobolev_classes}
-We now consider the situation where the regression function displays some higher-order regularity, $f_0 \in H_0^s(\mc{X})$.
-We show that Laplacian eigenmaps methods continue to be optimal for all orders of $s$, as long as the design density is itself also sufficiently regular, $p \in C^{s - 1}(\mc{X})$.
-In estimation, this is the case for any dimension $d$, whereas in testing it is the case only when $d \leq 4$.
-
-\paragraph{Estimation.}
-In order to show that $\wh{f}$ is an optimal estimator over $H_0^s(\mc{X};M)$, we will require that $\varepsilon$ be meaningfully larger than the lower bound in~\ref{asmp:parameters_estimation_fo}.
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{2}
-	\item 
-	\label{asmp:parameters_estimation_ho}
-	For constants $c_0$ and $C_0$, the graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy
-	\begin{equation}
-	\label{eqn:radius_ho}
-	C_0\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/d}, (M^2n)^{-1/(2(s - 1) + d)}\biggr\} \leq \varepsilon \leq c_0\min\{1, K^{-1/d}\}
-	\end{equation}
-	and
-	\begin{equation*}
-	K = \min\Bigl\{\floor{(M^2n)^{d/(2s + d)}} \vee 1,n\Bigr\}
-	\end{equation*}
-\end{enumerate}
-Crucially, when $n$ is sufficiently large the two conditions in~\ref{asmp:parameters_estimation_ho} are guaranteed to not be mutually exclusive.
-This is because so long as $M^2 = \omega(n^{-1})$ then $(M^2n)^{-2/(2(s - 1) + d)} = o((M^2n)^{-2/(2s + d)})$, regardless of $s$ and $d$.
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_estimation_ho}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H_0^s(\mc{X},M)$ and $p \in C^{s - 1}(\mc{X})$.
-There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_ho}, then
-	\begin{equation}
-	\label{eqn:laplacian_eigenmaps_estimation_ho}
-	\|\wh{f} - f_0\|_n^2 \leq C\Bigl(\frac{1}{\delta}M^2(M^2n)^{-2s/(2s + d)} \wedge 1\Bigr) \vee \frac{1}{n},
-	\end{equation}
-	with probability at least $1 - \delta - Cn\exp(-cn\varepsilon^d) - \exp(-K)$.
-\end{theorem}
-Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that in the flat Euclidean setting Laplacian eigenmaps is a minimax rate-optimal estimator over Sobolev classes, for all values of $s$ and $d$.
-Some other remarks:
-\begin{itemize}
-	\item \emph{Sub-critical Sobolev spaces}.
-Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and~\ref{thm:laplacian_eigenmaps_estimation_ho} do not require that the smoothness index $s$ of the Sobolev space satisfy $s > d/2$, a condition often seen in the literature.
-In the sub-critical regime $s \leq d/2$, the Sobolev space $H^s(\mc{X})$ is quite irregular.
-It is not a Reproducing Kernel Hilbert Space (RKHS), nor does it continuously embed into $C^0(\mc{X})$, much less into any H\"{o}lder space.
-As a result, for certain versions of the nonparametric regression problem---e.g.
-when loss is measured in $\Leb^{\infty}$ norm, or when the design points $\{X_1,\ldots,X_n\}$ are assumed to be fixed---in a minimax sense even consistent estimation is not possible.
-Likewise, certain estimators are ``off the table'', most notably RKHS-based methods such as thin-plate splines of degree $k \leq d/2$.
-Nevertheless, for random design regression with error measured in squared $\Leb^2(P)$-norm, the spectral projection estimator $\wt{f}$ obtains the standard minimax rates $n^{-2s/(2s + d)}$ for all values of $s$ and $d$.
-Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and \ref{thm:laplacian_eigenmaps_estimation_ho} show that the same is true with respect to Laplacian eigenmaps, when error is measured in squared $\Leb^2(P_n)$-norm.
-	\item \emph{Smoothness of design density}.
-As promised,  Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} shows that Laplacian eigenmaps achieves optimal rates of convergence so long as the unknown design density $p$ is sufficiently smooth, $p \in C^{s - 1}(\mc{X})$.
-The requirement $p \in C^{s - 1}(\mc{X})$ is essential to showing that $\wh{f}$ enjoys the faster minimax rates of convergence when $s > 1$,  as we discuss in Section~\ref{subsec:analysis}.
-	\item \emph{Computational considerations}.
-The lower bound on $\varepsilon$ in~\ref{asmp:parameters_estimation_ho} will result in a dense neighborhood graph $G$, meaning the average degree of $G$ will grow polynomially in the sample size $n$ as $n \to \infty$.
-As compared to a sparse $G$, this results in more non-zero entries in the graph Laplacian, and increases the computational burden involved in computing $\wh{f}$.
-To address this issue, in Appendix~\ref{sec:computational_considerations} we review some approaches to \emph{spectral sparsification}, in which one efficiently computes a sparse graph $\wc{G}$ that approximates $G$ in a spectral sense (see).
-Letting $\wc{f}$ be the Laplacian eigenmaps estimator computed with respect to the sparsified graph $\wc{G}$, we provide upper bounds on $\|\wc{f} - f_0\|_n^2$, which show that under mild conditions on $\wc{G}$---which provably achieved by many techniques for spectral sparsification---the estimator $\wc{f}$ achieves the same rates of convergence as $\wh{f}$.
-\end{itemize}
-
-\paragraph{Testing.} The test $\varphi$ can adapt to the higher-order smoothness of $f_0$, when $\varepsilon$ and $K$ are chosen correctly.
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{3}
-	\item 
-	\label{asmp:parameters_testing_ho}
-	The graph radius $\varepsilon$ satisfies~\eqref{eqn:radius_ho}, and the number of eigenvectors
-	\begin{equation}
-	\label{eqn:eigenvector_testing_ho}
-	K = \min\Bigl\{\floor{(M^2n)^{2d/(4s + d)}} \vee 1, n\Bigr\}.
-	\end{equation}
-\end{enumerate}
-When $d \leq 4$, for any value of $s \in \mathbb{N}$ when $n$ is sufficiently large it is possible to choose $\varepsilon$ and $K$ such that both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho} are satisfied, and our next theorem establishes that in this situation $\varphi$ is an optimal test.
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_testing_ho}
-	Fix $a,b \in (0,1)$.
-Suppose Model~\ref{def:model_flat_euclidean}.
-Then $\mathbb{E}_0[\varphi] \leq a$, i.e $\varphi$ is a level-$a$ test.
-Suppose additionally $f_0 \in H_0^s(\mc{X},M)$, that $p \in C^{s-1}(\mc{X})$, and that $d \leq 4$.
-Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n \geq N$: if the Laplacian eigenmaps test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_ho}, and if $f_0$ satisfies
-	\begin{equation}
-	\label{eqn:laplacian_eigenmaps_testing_criticalradius_ho}
-	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-4s/(4s + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{2s/d}} \biggr) \vee \frac{1}{n},
-	\end{equation}
-	then $\Ebb_{f_0}[1 - \phi] \leq b$.
-\end{theorem}
-Similarly to the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that when $M \asymp 1$, then $\varphi$ is a minimax rate-optimal test over $H_0^s(\mc{X})$.
-However, unlike the first-order case, when $4 < d < 4s$ the minimax testing rate over $H_0^s(\mc{X})$ is still on the order of $M^2(M^2n)^{-4s/(4s + d)}$; unfortunately, we can no longer claim that $\varphi$ is an optimal test in this regime.
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_testing_ho_suboptimal}
-	Under the same setup as Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, but with $4 < d < 4s$.
-If the Laplacian eigenmaps test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, number of eigenvectors $K$ satisfying~\eqref{eqn:eigenvector_testing_ho}, and $\varepsilon = (M^2n)^{-1/(2(s - 1) + d)}$, and if
-	\begin{equation}
-	\label{eqn:laplacian_eigenmaps_testing_criticalradius_ho_suboptimal}
-	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-2s/(2(s - 1) + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{2s/d}} \biggr) \vee \frac{1}{n},
-	\end{equation}
-	then $\Ebb_{f_0}[1 - \phi] \leq b$.
-\end{theorem}
-Focusing again on the special case where $M \asymp 1$, Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} says that $\varphi$ has have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2(s - 1) + d)}$ and $4 < d < 4s$.
-This is smaller than the estimation rate $n^{-2s/(2s + d)}$, but larger than the minimax squared critical radius $n^{-4s/(4s + d)}$.
-As a technical matter, the problem is that when $d > 4$ there do not exist any choices of $\varepsilon$ and $K$ which satisfy both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho}, and as a result we cannot optimally balance (our upper bounds on) testing bias and variance (defined momentarily in~\eqref{eqn:testing_biasvariance}).
-Although we suspect $\varphi$ is truly suboptimal when $d > 4$, the inequality in~\eqref{eqn:testing_biasvariance} gives only an upper bound on testing bias, and thus we cannot rule out that the test $\varphi$ is optimal for all $4 < d < 4s$.
-We leave the matter to future work.
-
-\subsection{Analysis}
-\label{subsec:analysis}
-
-We now outline the high-level strategy we follow when proving each of Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}.
-We analyze the estimation error of $\wh{f}$, and the testing error of $\wh{\varphi}$, by first conditioning on the design points $X_1,\ldots,X_n$ and deriving \emph{design-dependent} bias and variance terms.
-For estimation, we have that with probability at least $1 - \exp(-K)$,
-\begin{equation}
-\label{eqn:estimation_biasvariance}
-\|\wh{f} - f_0\|_n^2 \leq \underbrace{\frac{\dotp{L_{n,\varepsilon}^s f_0}{f_0}_n}{\lambda_{K}^s}}_{\textrm{bias}} + \underbrace{\frac{5K}{n} \vphantom{\frac{\dotp{L^s f_0}{f_0}_n}{\lambda_{K}^s}}}_{\textrm{variance}}.
-\end{equation}
-For testing, we have that $\varphi$ (which is a level-$a$ test by construction) also has small Type II Error, $\Ebb_{f_0}[1 - \phi] \leq b/2$, if 
-\begin{equation}
-\label{eqn:testing_biasvariance}
-\|f_0\|_n^2 \geq  \underbrace{\frac{\dotp{L_{n,\varepsilon}^s f_0}{f_0}_n}{\lambda_{K}^s}}_{\textrm{bias}} + \underbrace{32\frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]}_{\textrm{variance}}.
-\end{equation}
-These design-dependent bias-variance decompositions are reminiscent of the more classical bias-variance decompositions typical in the analysis of population-level spectral projection methods, but different in certain key respects.
-Comparing~\eqref{eqn:estimation_biasvariance} and~\eqref{eqn:testing_biasvariance} to~\eqref{pf:spectral_series_estimation_3} and~\eqref{pf:spectral_series_test}, we see that two continuum objects in the latter pair of bounds, the Sobolev norm $\|f_0\|_{\mc{H}^s(\mc{X})}^2$ and the eigenvalue $\lambda_k(\Delta_P)$, have been replaced by graph-based analogues: the graph Sobolev seminorm $\dotp{L_{n,\varepsilon}^sf_0}{f_0}_n$ and the graph Laplacian eigenvalue $\lambda_k^s$.
-These latter quantities, along with the empirical squared norm $\|f_0\|_n^2$, are random variables that depend on the random design points $X_1,\ldots,X_n$.
-We proceed to establish suitable upper and lower bounds that hold in probability.
-
-\paragraph{Estimates of graph quadratic forms.}
-In Proposition~\ref{prop:graph_seminorm_fo} we restate an upper bound on the Dirichlet energy $\dotp{L_{n,\varepsilon}f}{f}_n$ from~\cite{green2021}. 
-\begin{proposition}[Lemma 1 of~\cite{green2021}]
-	\label{prop:graph_seminorm_fo}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f \in H^1(\mc{X})$.
-There exist constants $c,C$ that do not depend on $f$ or $n$ such that the following statement holds for any $\delta \in (0,1)$: if $\eta$ satisfies~\ref{asmp:kernel} and $\varepsilon < c$, then
-	\begin{equation}
-	\label{eqn:graph_seminorm_fo}
-	\dotp{L_{n,\varepsilon}f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^1(\mc{X})}^2,
-	\end{equation}
-	with probability at least $1 - \delta$.
-\end{proposition}
-Proposition~\ref{prop:graph_seminorm_fo} follows by upper bounding the expectation of $\Ebb\dotp{L_{n,\varepsilon}f}{f}_n = \dotp{L_{P,\varepsilon}f}{f}_P$---where $L_{P,\varepsilon}$ is the non-local Laplacian operator defined in~\eqref{eqn:nonlocal_laplacian}---by (a constant times) the squared Sobolev norm $\|f\|_{H^1(\mc{X})}^2$, and an application of Markov's inequality.
-
-In this work, we establish that an analogous bound holds for $\dotp{L_{n,\varepsilon}^sf_0}{f_0}_n$ when $s > 1$. We call this quantity the \emph{graph Sobolev semi-norm}.
-\begin{proposition}
-	\label{prop:graph_seminorm_ho} 
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f \in H_0^s(\mc{X})$ and $p \in C^{s - 1}(\mc{X})$.
-Then there exist constants $c$ and $C$ that do not depend on $f_0$ or $n$ such that the following statement holds for any $\delta \in (0,1)$: if $\eta$ satisfies~\ref{asmp:kernel_flat_euclidean} and $Cn^{-1/(2(s - 1) + d)} < \varepsilon < c$, then
-	\begin{equation}
-	\label{eqn:graph_seminorm_ho}
-	\dotp{L_{n,\varepsilon}^s f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^s(\mc{X})}^2 ,
-	\end{equation}
-	with probability at least $1 - \delta$.
-\end{proposition}
-We now summarize the techniques used to prove Proposition~\ref{prop:graph_seminorm_ho}, which will help explain what role the conditions on $f_0$, $p$ and $\varepsilon$ play.
-To upper bound $\dotp{L_{n,\varepsilon}^sf}{f}_n$ in terms of $\|f\|_{H^s(\mc{X})}^2$, we introduce an intermediate quantity: the \emph{non-local Sobolev seminorm} $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$.
-This seminorm is defined with respect to the iterated non-local Laplacian $L_{P,\varepsilon}^s = L_{P,\varepsilon} \circ \cdots \circ L_{P,\varepsilon}$, where $L_{P,\varepsilon}$ is a non-local approximation to $\Delta_P$,
-\begin{equation}
-\label{eqn:nonlocal_laplacian}
-L_{P,\varepsilon}f(x) := \frac{1}{\varepsilon^{d + 2}}\int_{\mc{X}}\bigl(f(z) - f(x)\bigr) \eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dP(x).
-\end{equation}
-Then the proof of Proposition~\ref{prop:graph_seminorm_ho} proceeds according to the following steps.
-\begin{itemize}
-	\item First we note that $\dotp{L_{n,\varepsilon}^s f}{f}_n$ is itself a biased estimate of the non-local seminorm $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$.
-Specifically, $\dotp{L_{n,\varepsilon}^s f}{f}_n$ is a $V$-statistic, meaning it is the sum of an unbiased estimator of $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$ (in other words, a $U$-statistic) plus some higher-order, pure bias terms.
-We show that these pure bias terms are negligible when $\varepsilon = \omega(n^{-1/(2(s - 1) + d)})$.
-	\item For $x$ sufficiently far from the boundary of $\mc{X}$---precisely $x \in \mc{X}$ such that $B(x,j\varepsilon) \subseteq \mc{X}$---we show that $L_{P,\varepsilon}^jf(x) \to \sigma_{\eta}^j \Delta_P^jf(x)$ as $\varepsilon \to 0$.
-Here $j = (s - 1)/2$ when $s$ is odd and $j = (s - 2)/2$ when $s$ is even.
-This step bears some resemblance to the analysis of the bias term in kernel smoothing, and requires that $p \in C^{s-1}(\mc{X})$.
-	\item On the other hand for $x$ sufficiently near the boundary of $\mc{X}$, $L_{P,\varepsilon}^jf(x)$ does not in general converge to $\sigma_{\eta}^j\Delta_P^jf(x)$.
-Instead, we use the zero-trace property of $f$ to show that $L_{P,\varepsilon}^jf(x)$ is small.
-	\item Finally, we combine the results of previous two steps to deduce an upper bound on $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$ in terms of the squared Sobolev norm $\|f\|_{H^s(\mc{X})}^2$.
-Roughly speaking, when $s$ is odd, $\dotp{L_{P,\varepsilon}^sf}{f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}^jf) \approx \sigma_{\eta}^{2j}E_{P,\varepsilon}(\Delta_P^jf)$, whereas when $s$ is even $\dotp{L_{P,\varepsilon}^sf}{f}_P = \|L_{P,\varepsilon}L_{P,\varepsilon}^{j}f\|_{P}^2 \approx \sigma_{\eta}^{2j}\|L_{P,\varepsilon} \Delta_Pf\|_P^2$.
-Reasoning in this way, we can translate estimates of $L_{P,\varepsilon}^jf$ into an upper bound on the order-$s$ non-local Sobolev seminorm, even though $s > j$.
-\end{itemize}
-Together, these steps establish Proposition~\ref{prop:graph_seminorm_ho}.
-It is worth pointing out that we do not try to establish the pointwise estimate $L_{P,\varepsilon}^sf \to \sigma_{\eta}^s\Delta_{P}^sf$ in $L^2(P)$.
-If we had such an estimate, it would immediately follow that $\dotp{L_{P,\varepsilon}^sf}{f}_{P} \to \sigma_{\eta}^s\dotp{\Delta_P^sf}{f}_{P}$.
-Unfortunately, we assume only that $f$ has $s$ bounded derivatives, while $\Delta_P^s$ is an order-$2s$ differential operator; thus in general $L_{P,\varepsilon}^sf$ may not approach $\sigma_{\eta}^s\Delta_{P}^sf$ as $\varepsilon \to 0$.
-Instead we opt for the slightly more complicated approach outlined above, in which we only ever need show that $L_{P,\varepsilon}^jf(x) \to \sigma_{\eta}^j \Delta_P^jf(x)$ for some $j < s/2$.
-
-\paragraph{Neighborhood graph eigenvalues.}
-On the other hand, several recent works \citep{burago2014,garciatrillos18,calder2019} have analyzed the convergence of $\lambda_{k}$ towards $\lambda_{k}(\Delta_P)$.
-They provide explicit bounds on the relative error $|\lambda_{k} - \lambda_{k}(\Delta_P)|/\lambda_{k}(\Delta_P)$, which show that the relative error is small for sufficiently large $n$ and small $\varepsilon$.
-Crucially, the guarantees hold simultaneously for all $1 \leq k \leq K$ as long as $\lambda_{K}(\Delta_P) = O(\varepsilon^{-2})$.
-These results are actually stronger than are necessary to establish Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho}---in order to get rate-optimality, we need only show that for the relevant values of $K$, $\lambda_{K}/\lambda_K(P) = \Omega_P(1)$---but unfortunately they all assume $P$ is supported on a manifold without boundary (i.e.
-they assume Model~\ref{def:model_manifold} rather than Model~\ref{def:model_flat_euclidean}).
-
-In the case where $\mc{X}$ is assumed to have a boundary, the graph Laplacian $L_{n,\varepsilon}$ is a reasonable approximation of the operator $\Delta_P$ at points $x \in \mc{X}$ for which $B(x,\varepsilon) \subseteq \mc{X}$.
-In contrast, at points $x$ near the boundary of $\mc{X}$, the graph Laplacian is known to approximate a different operator altogether \citep{belkin2012}.\footnote{This is directly related to the boundary bias of kernel smoothing, since the graph Laplacian can be viewed as a kernel-based estimator of $\Delta_P$.} This renders analysis of $\lambda_k$ substantially more challenging, since its continuum limit is not $\lambda_k(\Delta_P)$.
-Rather than establishing the convergence of $\lambda_k$, we will instead use Lemma~2 of \cite{green2021}, whose assumptions match our own, and who give a weaker bound on the ratio $\lambda_k/\lambda_k(\Delta_P)$ that will nevertheless suffice for our purposes.
-
-\begin{proposition}[Lemma~2 of \cite{green2021}]
-	\label{prop:graph_eigenvalue}
-	Suppose Model~\ref{def:model_flat_euclidean}.
-Then there exist constants $c$ and $C$ such that the following statement holds: if $\eta$ satisfies~\ref{asmp:kernel_flat_euclidean} and $C(\log n/n)^{1/d} < \varepsilon < c$, then
-	\begin{equation}
-	\label{eqn:graph_eigenvalue}
-	\lambda_k \geq c \cdot \min\Bigl\{\lambda_k(\Delta_P), \frac{1}{\varepsilon^{2}} \Bigr\} \quad \textrm{for all $1 \leq k \leq n$,}
-	\end{equation}
-	with probability at least $1 - Cn\exp\{-c n\varepsilon^d\}$. 
-\end{proposition}
-By our assumptions on $P$, $\lambda_0(\Delta_P) = \lambda_0 = 0$.
-Furthermore, Weyl's Law~\eqref{eqn:weyl} tells us that under Model~\ref{def:model_flat_euclidean}, $k^{2/d} \lesssim \lambda_{k}(\Delta_P) \lesssim k^{2/d}$ for all $k \in \mathbb{N}, k > 1$.
-Combining these statements with~\eqref{eqn:graph_eigenvalue}, we conclude that $\lambda_{K} = \Omega_P(K^{2/d})$ so long as $K \lesssim \varepsilon^{-d}$.
-
-\paragraph{Empirical norm.}
-Finally, in Proposition~\ref{prop:empirical_norm_sobolev} we establish that a one-sided bound of the form $\|f_0\|_n^2 \gtrsim \|f_0\|_P^2$ whenever $\|f_0\|_P^2$ is itself sufficiently large.
-\begin{proposition}
-	\label{prop:empirical_norm_sobolev}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f \in H^s(\Xset,M)$ for some $s > d/4$.
-There exist constants $c$ and $C$ that do not depend on $f_0$ or $n$ such that the following statement holds for any $\delta > 0$:  if
-	\begin{equation}
-	\label{eqn:empirical_norm_sobolev_1}
-	\norm{f}_{P} \geq C M \biggl(\frac{1}{\delta n}\biggr)^{s/d}
-	\end{equation}
-	then with probability at least $1 - \exp\{-(cn \wedge 1/\delta)\}$,
-	\begin{equation}
-	\label{eqn:empirical_norm_sobolev}
-	\norm{f}_n^2 \geq \frac{1}{2} \|f_0\|_P^2.
-	\end{equation}
-\end{proposition}
-To prove Proposition~\ref{prop:empirical_norm_sobolev}, we use a Gagliardo-Nirenberg interpolation inequality (see e.g.
-Theorem~12.83 of~\citep{leoni2017}) to control the $4$th moment of $f \in H^s(\mc{X})$ in terms of $\|f\|_P$ and $|f|_{H^s(\mc{X})}$, then invoke a one-sided Bernstein's inequality as in \cite[Section 14.2]{wainwright2019}.
-Note carefully that the statement~\eqref{eqn:empirical_norm_sobolev} is \emph{not} a uniform guarantee over all $f \in H^s(\mc{X};M)$.
-Indeed, such a statement cannot hold in the sub-critical regime ($2s \leq d$).\footnote{This is because in the sub-critical regime, for any set of points $\{x_1,\ldots,x_n\}$ there exist a sequence of functions $(f_k)_{k \in \mathbb{N}} \subset H^s(\mc{X};1)$ satisfying $f_k(x_i) = 1$ for each $i = 1,\ldots,n$, but for which $\|f_k\|_P^2 \to 0$ as $k \to \infty$.} Fortunately, a pointwise bound---meaning a bound that holds with high probability for a single $f \in H^s(\mc{X})$---is sufficient for our purposes.
-
-Finally, invoking the bounds of Propositions~\ref{prop:graph_seminorm_fo}-\ref{prop:empirical_norm_sobolev} inside the bias-variance tradeoffs~\eqref{eqn:estimation_biasvariance} and~\eqref{eqn:testing_biasvariance} and then choosing $K$ to balance bias and variance (when possible), leads to the conclusions of Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}.
-
-% (AG 8/29/21): Following chunk commented out but not deleted, in case it proves useful at a later point.
-% We conclude this section by pointing out that, at first glance, it may be somewhat remarkable that Laplacian eigenmaps can ever attain optimal rates of convergence.
-The sharpest known results \citep{calder2019,cheng2021} show that the squared error with which a graph Laplacian eigenvector approximates its continuum limit, $\|v_k - \psi_k\|_n^2$,  converges to $0$ at a rate of $C_k n^{-2/(4 + d)}$, for a constant $C_k \to \infty$ as $k \to \infty$.
-For any fixed $k$, this is slower than the optimal estimation or testing rates, and when the number of eigenvectors $K$ increases with $n$, as is necessary to optimally balance bias and variance of Laplacian eigenmaps, the issue only gets worse.
-Moreover, there are reasons to believe that this upper bound on $\|v_k - \psi_k\|_n^2$ is unimprovable (see~\cite{cheng2021}).
-If so, this suggests that Laplacian eigenmaps can learn regression function more efficiently, from a statistical perspective, than Laplacian eigenvectors can learn population-level eigenfunctions.
-Put more simply: regression is a forgiving problem.
-
-\section{Manifold Adaptivity}
-\label{sec:manifold_adaptivity}
-
-In this section we consider the manifold setting, where $(X_1,Y_1),\ldots,(X_n,Y_n)$ are observed according to Model~\ref{def:model_manifold}.
-In this setting, it is known that the minimax rates depend only on the intrinsic dimension $m$; more specifically,~\cite{bickel2007,ariascastro2018} show that for H\"{o}lder classes, the minimax estimation rate is $n^{-2s/(2s + m)}$ and the testing rate is $n^{-4s/(4s + m)}$\footnote{Although~\cite{ariascastro2018} considers density testing, usual arguments regarding equivalence of experiments~\cite{brown96} imply that the same rates apply to regression testing}.
-On the other hand, a theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear.
-We build on this work by showing that when $f_0 \in H^s(\mc{X})$ and $P$ is supported on a manifold, Laplacian eigenmaps achieve the sharper minimax estimation and testing rates.
-
-\subsection{Laplacian eigenmaps error rates under the manifold hypothesis}
-Unlike in the flat-Euclidean case, since Model~\ref{def:model_manifold} assumes that $\mc{X}$ is boundaryless it is easy to deal with the first-order $(s = 1)$ and higher-order $(s > 1)$ cases all at once.
-A more important distinction between the results of this section and those of Section~\ref{sec:minimax_optimal_laplacian_eigenmaps} is that we will establish Laplacian eigenmaps is optimal only when the regression function $f_0 \in H^s(\mc{X};M)$ for $s \in \{1,2,3\}$.
-Otherwise, this section will proceed in a similar fashion to Section~\ref{sec:higher_order_sobolev_classes}.
-
-\paragraph{Estimation.}
-To ensure that $\wh{f}$ is an in-sample minimax rate-optimal estimator, we choose the kernel function $\eta$, graph radius $\varepsilon$ and number of eigenvectors $K$ as in~\ref{asmp:parameters_estimation_ho}, except with ambient dimension $d$ replaced by the intrinsic dimension $m$.
-
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{4}
-	\item 
-	\label{asmp:kernel_manifold}
-	The kernel function $\eta$ is a nonincreasing function supported on a subset of $[0,1]$. Its restriction to $[0,1]$ is Lipschitz, and $\eta(1/2) > 0$. Additionally, it is normalized so that
-	\begin{equation*}
-	\int_{\Reals^m} \eta(\|z\|) \,dz = 1,
-	\end{equation*}
-	and we assume \smash{$\int_{\Reals^m} \|x\|^2 \eta(\|x\|) \,dx < \infty$}.
-	\item 
-	\label{asmp:parameters_estimation_manifold}
-	For a constant $C_0$, the graph radius $\varepsilon$ and number of eigenvectors $K$ satisfy
-	\begin{equation}
-	\label{eqn:radius_estimation_manifold}
-	C_0\max\biggl\{\biggl(\frac{\log}{n}\biggr)^{1/m}, n^{-1/(2(s - 1) + m)}\biggr\} \leq \varepsilon \leq \min\{\mathrm{inj}(\mc{X}), K^{-1/m}\},
-	\end{equation}
-	where we recall that $\mathrm{inj}(\mc{X})$ is a lower bound on the injectivity radius of $\mc{X}$. Additionally,
-	\begin{equation*}
-	K = \min\Bigl\{\floor{(M^2n)^{m/(2s + m)}} \wedge 1,n \Bigr\}.
-	\end{equation*}
-\end{enumerate}
-
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_estimation_manifold}
-	Suppose Model~\ref{def:model_manifold}, and additionally $f_0 \in H^s(\mc{X},M)$ and $p \in C^{s - 1}(\mc{X})$ for $s \leq 3$.
-There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_manifold}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_manifold}, then
-	\begin{equation}
-	\label{eqn:laplacian_eigenmaps_estimation_manifold}
-	\|\wh{f} - f_0\|_n^2 \leq C\Bigl(\frac{1}{\delta}M^2(M^2n)^{-2s/(2s + m)} \wedge 1\Bigr) \vee \frac{1}{n},
-	\end{equation}
-	with probability at least $1 - \delta - Cn\exp(-cn\varepsilon^m) - \exp(-K)$.
-\end{theorem}
-
-\paragraph{Testing.}
-Likewise, to construct a minimax optimal test using $\wh{T}$, we choose $\varepsilon$ and $K$ as in~\ref{asmp:parameters_testing_fo}, except with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{5}
-	\item 
-	\label{asmp:parameters_testing_manifold}
-	The graph radius $\varepsilon$ satisfies~\eqref{eqn:radius_estimation_manifold}, and the number of eigenvectors
-	\begin{equation*}
-	K = \min\Bigl\{\floor{(M^2n)^{2m/(4s + m)}} \wedge 1,n \Bigr\}.
-	\end{equation*}
-\end{enumerate}
-
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_testing_manifold}
-	Fix $a,b \in (0,1)$.
-Suppose Model~\ref{def:model_manifold}.
-Then $\mathbb{E}_0[\varphi] \leq a$, i.e $\varphi$ is a level-$a$ test.
-Suppose additionally $f_0 \in H^s(\mc{X},M)$, that $p \in C^{s-1}(\mc{X})$, and that $s \leq 3$ and $m \leq 4$.
-Then there exist constants $c$, $C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_manifold}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_manifold}, and if $f_0$ satisfies
-	\begin{equation}
-	\label{eqn:laplacian_eigenmaps_testing_criticalradius_manifold}
-	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-4s/(4s + m) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{2s/m}} \biggr) \vee \frac{1}{n},
-	\end{equation}
-	then $\Ebb_{f_0}[1 - \phi] \leq b$.
-\end{theorem}
-
-Unlike in the full-dimensional case, in the manifold setting our upper bounds on the estimation and testing error of Laplacian eigenmaps match the minimax rate only when $s \leq 3$.
-When $s \geq 4$, the containment $H^s(\mc{X};1) \subset H^{3}(\mc{X};1)$ (taking the radius of the Sobolev ball $M = 1$ for simplicity) implies that the Laplacian eigenmaps estimator $\wh{f}$ has in-sample mean-squared error of at most $n^{-6/(6 + m)}$, and that the Laplacian eigenmaps test has small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-12/(12 + d)}$; however, these are slower than the usual minimax rates.
-
-We now explain this difference between the flat Euclidean and manifold settings.
-At a high level, thinking of the graph $G$ as an estimate of the manifold $\mc{X}$, we incur some error by using Euclidean distance rather than geodesic distance to form the edges of $G$.
-This is in contrast with the full-dimensional setting, where the Euclidean metric exactly coincides with the geodesic distance for all points $x,z \in \mc{X}$ that are sufficiently close to each other and far from the boundary of $\mc{X}$.
-This extra error incurred in the manifold setting by using the ``wrong distance'' dominates when $s \geq 4$.
-
-As this explanation suggests, by building $G$ using the geodesic distance one could avoid this error, and might obtain superior rates of convergence.
-However this is not an option for us, as we assume $\mc{X}$---and in particular its geodesics---are unknown.
-Likewise, a classical spectral projection estimator, using eigenfunctions of the manifold Laplace-Beltrami operator, will achieve the minimax rate for all values of $s$ and $m$; but this is undesirable for the same reason---we do not want to assume that $\mc{X}$ is known.
-It is not clear whether this gap between spectral projection and Laplacian eigenmaps estimators---or more generally, between estimators which assume the manifold is known, and those which do not---is real, or a product of loose upper bounds.
-
-Finally, as in the full-dimensional case, when the intrinsic dimension $m > 4$ we cannot choose the graph radius $\varepsilon$ and number of eigenvectors $K$ to optimally balance bias and variance.
-Instead, reasoning as in the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} shows that when $1 \leq s \leq 3$, the Laplacian eigenmaps test has critical radius as given by~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_ho_suboptimal}, but with the ambient dimension $d$ replaced by $m$.
-% (SB via AG): Maybe some comment on the difference between testing and estimation.
-
-\paragraph{Analysis.}
-The high-level strategy used to prove Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold} is the same as in the flat-Euclidean setting.
-More specifically, we will use precisely the same bias-variance decompositions~\eqref{eqn:estimation_biasvariance} (for estimation) and~\eqref{eqn:testing_biasvariance} (for testing).
-The difference will be that our bounds on the graph Sobolev seminorm $\dotp{L_{n,\varepsilon}^sf_0}{f_0}_n$, graph eigenvalue $\lambda_K$, and empirical norm $\|f_0\|_n^2$ will now always depend on the intrinsic dimension $m$, rather than the ambient dimension $d$.
-The precise results we use are contained in Propositions~\ref{prop:graph_seminorm_manifold}-\ref{prop:empirical_norm_sobolev_manifold}.
-\begin{proposition}
-	\label{prop:graph_seminorm_manifold} 
-	Suppose Model~\ref{def:model_manifold}, and additionally that $f_0 \in H^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$ for $s = 1,2$ or $3$.
-Then there exist constants $c_0,C_0$ and $C$ that do not depend on $f_0$, $n$ or $M$ such that the following statement holds for any $\delta \in (0,1)$: if $\eta$ satisfies~\ref{asmp:kernel_manifold} and $C_0n^{-1/(2(s - 1) + m)} < \varepsilon < c_0$, then
-	\begin{equation}
-	\label{eqn:graph_seminorm_manifold}
-	\dotp{L_{n,\varepsilon}^s f}{f}_n \leq \frac{C}{\delta} \|f\|_{H^s(\mc{X})}^2,
-	\end{equation}
-	with probability at least $1 - 2\delta$.
-\end{proposition}
-
-As discussed previously, when $\mc{X}$ is a domain without boundary and $\Delta_P$ is the manifold weighted Laplace-Beltrami operator, appropriate bounds on the graph eigenvalues $\lambda_k$ have already been derived in \citep{burago2014,trillos2019,garciatrillos19}.
-The precise result we need is a direct consequence of Theorem 2.4 of~\citep{calder2019}.
-\begin{proposition}[\textbf{c.f Theorem 2.4 of~\citep{calder2019}}]
-	\label{prop:graph_eigenvalue_manifold}
-	Suppose Model~\ref{def:model_manifold}.
-Then there exist constants $c$ and $C$ such that the following statement holds: if $\eta$ satisfies~\ref{asmp:kernel_manifold} and $C(\log n/n)^{1/m} < \varepsilon < c$, then
-	\begin{equation}
-	\label{eqn:graph_eigenvalue_manifold}
-	\lambda_k \geq c \cdot \min\Bigl\{\lambda_k(\Delta_P), \frac{1}{\varepsilon^{2}} \Bigr\} \quad \textrm{for all $1 \leq k \leq n$,}
-	\end{equation}
-	with probability at least $1 - Cn\exp\{-c n\varepsilon^d\}$. 
-\end{proposition}
-(For the specific computation used to deduce Proposition~\ref{prop:graph_eigenvalue_manifold} from Theorem 2.4 of~\citep{calder2019}, see~\cite{green2021}.)
-
-Finally, we have the following lower bound on the empirical norm $\|f\|_n$ under the hypotheses of Model~\ref{def:model_manifold}. 
-\begin{proposition}
-	\label{prop:empirical_norm_sobolev_manifold}
-	Suppose Model~\ref{def:model_manifold}, and additionally that $f_0 \in H^s(\Xset,M)$ for some $s > m/4$.
-There exists a constant $C$ that does not depend on $f_0$ such that the following statement holds for all $\delta > 0$:  if
-	\begin{equation}
-	\label{eqn:empirical_norm_sobolev_manifold_1}
-	\norm{f_0}_{P} \geq \frac{C M}{\delta^{s/m}}n^{-s/m},
-	\end{equation}
-	then with probability at least $1 - \exp\{-(cn \wedge 1/\delta)\}$,
-	\begin{equation}
-	\label{eqn:empirical_norm_sobolev_manifold}
-	\norm{f_0}_n^2 \geq \frac{1}{2} \|f_0\|_P^2.
-	\end{equation}
-\end{proposition}
-We prove Proposition~\ref{prop:empirical_norm_sobolev_manifold} in a parallel manner to its flat Euclidean counterpart (Proposition~\ref{prop:empirical_norm_sobolev}), by first using a Gagliardo-Nirenberg inequality to upper bound the $L^4(\mc{X})$ norm of a Sobolev function defined on a compact Riemannian manifold, and then applying a one-sided Bernstein's inequality.
-Finally, combining Propositions~\ref{prop:graph_seminorm_manifold}-\ref{prop:empirical_norm_sobolev_manifold} with the conditional-on-design bias-variance decompositions~\eqref{eqn:estimation_biasvariance} and \eqref{eqn:testing_biasvariance} leads to the conclusions of Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold}.
-
-\section{Out-of-sample error}
-\label{sec:out_of_sample}
-Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity} show that $\wh{f}$ is a minimax optimal estimator over Sobolev spaces, when the loss is  \emph{in-sample}---that is, measured in $\Leb^2(P_n)$ norm.
-In this section we consider \emph{out-of-sample} error---error measured in $L^2(P)$ norm---which is the more typical metric used in the study of nonparametric regression with random design.
-
-Of course, the Laplacian eigenmaps estimator is only defined at the design points $\{X_1,\ldots,X_n\}$, and to measure error in $L^2(P)$ norm we must first extend it to be defined over the entire domain $\Xset$.
-We propose a simple method, kernel smoothing, to do the job.
-The method can applied to any estimator defined at the design points, including Laplacian eigenmaps, and we show that a smoothed version of our original estimator $\wh{f}$ has optimal $L^2(P)$ error.
-For simplicity, in this section we will stick to the flat Euclidean setting, where $(X_1,Y_1),\ldots,(X_n,Y_n)$ are observed according to Model~\ref{def:model_flat_euclidean}.
-
-\paragraph{Extension by kernel smoothing.}
-We now formally define our approach to extension by kernel smoothing.
-For a kernel function $\psi(\cdot): [0,\infty) \to (-\infty,+\infty)$, bandwidth $h > 0$, and a distribution $Q$, the \emph{Nadaraya-Watson kernel smoother} $T_{Q,h}$ is given by
-\begin{equation*}
-\bigl(T_{Q,h}f)(x) := 
-\begin{dcases*}
-\frac{1}{d_{Q,h}(x)} \int_{\Omega} f(z)\psi\biggl(\frac{\|z - x\|}{h}\biggr) \,dQ(z), & \textrm{if $d_{Q,h}(x) > 0$,} \\
-0, &\textrm{otherwise,}
-\end{dcases*}
-\end{equation*}
-where $d_{Q,h}(x) := \int_{\Omega} \psi\bigl(\|z - x\|/h\bigr) \,dQ(z)$.
-For convenience, we will write $T_{n,h}f(x) := T_{P_n,h}f(x)$, and $d_{n,h}(x) := n \cdot d_{P_n,h}(x)$.
-We extend the Laplacian eigenmaps estimator by passing the kernel smoother $T_{n,h}$ over it, that is we consider the estimator $T_{n,h}\wh{f}$, which is defined at every $x \in \mc{X}$ (indeed, at every $x \in \Rd$).
-Note that ``extension'' here is a slight abuse of nomenclature, since $T_{n,h}\wh{f}$ and $\wh{f}$ may not agree in-sample.
-
-\paragraph{Out-of-sample error of kernel smoothed Laplacian eigenmaps.}
-
-In Lemma~\ref{lem:kernel_smoothing_insample}, we consider an arbitrary estimator $\wc{f} \in \Reals^n$.
-We show that the out-of-sample error $\|T_{n,h}\wc{f} - f_0\|_P^2$ can be upper bounded by three terms--- (a constant times) the in-sample error $\|\wc{f} - f_0\|_n^2$, and variance and bias terms that arise naturally in the analysis of kernel smoothing over noiseless data.
-We shall assume the following conditions on $\psi$ and $h$.
-\begin{enumerate}[label=(K\arabic*)]
-	\setcounter{enumi}{2}
-	\item
-	\label{asmp:kernel}
-	The kernel function $\psi$ is supported on a subset of $[0,1]$. Additionally, $\psi$ is Lipschitz continuous on $[0,1]$, and is normalized so that
-	\begin{equation*}
-	\int_{-\infty}^{\infty} \psi(|z|) \,dz = 1.
-	\end{equation*}
-\end{enumerate}
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{6}
-	\item
-	\label{asmp:bandwidth}
-	For constants $c_0$ and $C_0$, the bandwidth parameter $h$ satisfies
-	\begin{equation*}
-	C_0\biggl(\frac{\log(1/h)}{n}\biggr)^{1/d} \leq h \leq c_0.
-	\end{equation*}
-\end{enumerate}
-\begin{lemma}
-	\label{lem:kernel_smoothing_insample}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f_0 \in H^1(\mc{X};M)$ and $p \in C^1(\mc{X})$.
-If the kernel smoothing estimator $T_{n,h}\wc{f}$ is computed with a kernel $\psi$ satisfying~\ref{asmp:kernel} and bandwidth $h$ satisfying~\ref{asmp:bandwidth}, it holds that
-	\begin{equation}
-	\label{eqn:kernel_smoothing_insample}
-	\|T_{n,h}\wc{f} - f_0\|_P^2 \leq C\biggl(\|\wc{f} - f_0\|_n^2 + \frac{1}{\delta} \cdot \frac{h^2}{nh^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{P,h}f_0 - f_0\|_P^2\biggr),
-	\end{equation}
-	with probability at least $1 - \delta - Ch^d\exp\{-Cnh^d\}$. 
-\end{lemma}
-Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data, by a factor of $h^2$.
-On the other hand the bias term is typical, and when $\psi$ is an order-$s$ kernel a standard analysis (see Lemma~\ref{lem:kernel_smoothing_bias}) shows that $\|T_{P,h}f_0 - f_0\|_P^2 \lesssim |f|_{H^s(\mc{X})}^2 h^{2s}$.
-\begin{enumerate}[label=(K\arabic*)]
-	\setcounter{enumi}{3}
-	\item
-	\label{asmp:ho_kernel}
-	The kernel function $\psi$ is an order-$s$ kernel, meaning it satisfies
-	\begin{equation*}
-	\int_{-\infty}^{\infty} \psi(|z|) \,dz = 1, \quad \int_{-\infty}^{\infty} z^j \psi(|z|) \,dz = 0 ~~\textrm{for}~j = 1,\ldots, s + d - 2, \quad \textrm{and}~ \int_{-\infty}^{\infty} z^{s + d - 1} \psi(|z|) \,dz < \infty.
-	\end{equation*}
-\end{enumerate}
-% (SB via AG): Give an example of a kernel which satisfies these requirements.
-
-Choosing $h \asymp n^{-1/(2(s - 1) + d)}$ balances the kernel smoothing bias and variance terms in~\eqref{eqn:kernel_smoothing_insample}, and implies that for $f_0 \in H^s(\mc{X};M)$,
-\begin{equation}
-\label{eqn:kernel_smoothing_insample2}
-\|T_{n,h}\wc{f} - f_0\|_P^2 \leq C\biggl(\|\wc{f} - f_0\|_n^2 + \frac{M^2}{\delta}n^{-2s/(2(s - 1) + d)}\biggr).
-\end{equation}
-This analysis tells us that the additional error incurred by passing a kernel smoother over an in-sample estimator $\wc{f}$ is negligible compared to the minimax rate of estimation.
-Consequently, if $\wc{f}$ converges at the minimax rate in squared $L^2(P_n)$-norm, then $T_{n,h}\wc{f}$ will converge at the minimax rate in squared $L^2(P)$-norm.
-It follows immediately from Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} (when $s = 1$) or Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} (when $s > 1$), that $T_{n,h}\wh{f}$ achieves the optimal rate of convergence in $L^2(P)$.
-
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_estimation_out_of_sample}
-	Suppose Model~\ref{def:model_flat_euclidean}.
-There exist constants $c$, $C$, and $N$ that do not depend on $f_0$ or $n$ such that each the following statements hold with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\} - Ch^d\exp\{-cnh^d\}$,  for all $n \geq N$ and for any $\delta \in (0,1)$.
-	\begin{itemize}
-		\item If $f_0 \in H^1(\mc{X};M)$, the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_fo}, and the out-of-sample extension $T_{n,h}\wh{f}$ is computed with bandwidth $h = n^{-1/d}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel}, then
-		\begin{equation*}
-		\|T_{n,h}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2/(2 + d)}.
-		\end{equation*}
-		\item If $f_0 \in H_0^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$ for some $s \in \mathbb{N}, s > 1$, and the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_ho}, and the out-of-sample extension $T_{n,h}\wh{f}$ is computed with bandwidth $h = n^{-1/(2(s - 1) + d)}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel} and~\ref{asmp:ho_kernel}, then
-		\begin{equation*}
-		\|T_{n,h}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2s/(2s + d)}.
-		\end{equation*}
-	\end{itemize}
-\end{theorem}
-Some remarks:
-\begin{itemize}
-	\item \emph{Data-driven parameter tuning}.
-Since $T_{n,h}\wh{f}$ is defined out-of-sample, we can use sample splitting or cross validation methods to tune hyperparameters, which we could not do for the original estimator $\wh{f}$.
-For instance, we can (i) split the sample into two halves, (ii) use the first half to compute $T_{n,h}\wh{f}$ for various values of $\varepsilon$, $h$, and $K$, (iii) choose the optimal values of these three hyperparameters by minimizing error on the held out set.
-Practically speaking, cross-validation is one of the most common approaches to choosing hyperparameters.
-Theoretically, it is known that choosing hyper-parameters through sample splitting can result in estimators that optimally adapt to the order of regularity $s$ \citep{gyorfi2006}.
-In other words, it leads to estimators that are rate-optimal (up to $\log n$ factors), even when $s$ is unknown.
-We believe similar arguments should imply that $T_{n,h}\wh{f}$ is adaptive in this sense when $\varepsilon,h$ and $K$ are chosen by sample splitting, but the details should be worked out carefully.
-	\item \emph{Alternatives to extension by kernel smoothing}.
-There exist other approaches to extending a function $f$ from its evaluations $\{f(X_1),\ldots,f(X_n)\}$: for instance, minimum-norm interpolation in some RKHS or Banach space or 1-nearest neighbors regression.
-We consider extension by kernel smoothing because it is a simple and statistically optimal procedure that does not require any knowledge of the domain $\mc{X}$ or distribution $P$---as we have argued, this latter property is one of the main selling points of Laplacian eigenmaps as a tool for nonparametric regression.
-One potential drawback to kernel smoothing is that it can change the value of the estimate at data, meaning $\wh{f} \neq (\wc{f}(X_1),\ldots, \wc{f}(X_n))$.
-More sophisticated procedures based on Nystr\"{o}m extension of the eigenvectors $v_1,\ldots,v_K$~\cite{fowlkes2004} would inherit the generality of kernel smoothing while being genuine interpolators; however, their properties are substantially more difficult to analyze.
-
-	%Finally, the Nystr\"{o}m method extends  eigenvectors $v_k$ to functions $f_k: \mc{X} \to \Reals$ as follows,
-	%\begin{equation*}
-	%L_{n,\varepsilon}f_k = \lambda_k f_k~~\textrm{on $\mc{X}$, such that}~~f_k(X_i) = %v_k(X_i)~~\textrm{for $i = 1,\ldots,n$.}
-	%\end{equation*}
-	%Here $L_{n,\varepsilon}$ should be thought of as an operator acting on functions $f \in C(\mc{X})$ as follows,
-	%\begin{equation*}
-	%L_{n,\varepsilon}f(x) = \sum_{i = 1}^{n} (f(x) - f(X_i))\eta\Bigl(\frac{\|X_i - x\|}{\varepsilon}\Bigr)
-	%\end{equation*}
-	% The Nystr\"{o}m approach to extension makes intriguing (re-)use of the graph Laplacian $L_{n,\varepsilon}$ and its eigenvectors $v_k$.
-Similar methods have been analyzed in the context of kernel ridge regression (references), but little is known about its application to graph Laplacians.
-\end{itemize}
-
-\section{Experiments}
-\label{sec:experiments}
-
-In this section we empirically demonstrate that Laplacian Eigenmaps is a reasonably good alternative to spectral projection, even when $n$ is only moderately large.
-In order to compare the two methods, in our experiments we stick to the simple case where $P$ is the uniform distribution over $\mc{X} = [0,1]^d$, and we have simple closed-form expressions for the eigenfunctions of $\Delta_P$.
-In general, it is not easy to analytically compute these eigenfunctions, which is part of the appeal of Laplacian Eigenmaps.
-
-\begin{figure*}[tb]
-	\includegraphics[width=.245\textwidth]{figures/mse/mse_by_sample_size_1d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/mse/mse_by_sample_size_1d_2s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/mse/mse_by_sample_size_2d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/mse/mse_by_sample_size_2d_2s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_1d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_1d_2s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_2d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/out_of_sample_mse/mse_by_sample_size_2d_2s.pdf}
-	\caption{Mean squared error (mse) of Laplacian eigenmaps and spectral projection estimators.
-Top row: in-sample mse of Laplacian eigenmaps (\texttt{LE}) and a spectral projection estimator (\texttt{SP}) as a function of sample size $n$.
-Bottom row: out-of-sample mse of Laplacian eigenmaps plus kernel smoothing (\texttt{LE+KS}) and a spectral projection estimator.
-Each plot is on the log-log scale, and the results are averaged over 400 repetitions.
-All estimators are tuned for optimal average mse.
-The black line shows the minimax rate (in slope only; the intercept is chosen to match the observed error).}
-	\label{fig:fig1}
-\end{figure*}
-
-\paragraph{Estimation.}
-In our first experiment, we compare the mean-squared error of the Laplacian eigenmaps estimator $\wh{f}$  to that of its population-level counterpart $\wt{f}$.
-We vary the sample size from $n = 1000$ to $n = 4000$; sample $n$ design points $X_1,\ldots,X_n$ from the uniform distribution on the cube $[-1,1]^d$; and sample responses $Y_i$ according to~\eqref{eqn:model} with regression function $f_0 = M/\lambda_K^{s/2} \cdot \psi_K$ for $K \asymp n^{d/(2s + d)}$ (the pre-factor $M/\lambda_K^{s/2}$ is chosen so that $|f_0|_{H^s(\mc{X})}^2 = M^2$).
-In Figure~\ref{fig:fig1} we show the in-sample mean-squared error of the two estimators as a function of $n$, for different dimensions $d$ and order of smoothness $s$.
-We see that all estimators have mean-squared error converging to zero at roughly the minimax rate.
-While unsurprisingly population-level spectral projection has smaller error, generally speaking that the error of Laplacian Eigenmaps gets closer to that of population-level spectral projection as $n$ gets larger.
-
-We also compare the out-of-sample estimation error, over a held-out test set, of Laplacian eigenmaps plus kernel smoothing to the spectral projection estimator.
-The story is much the same, supporting our theoretical claim that the additional error incurred by kernel smoothing of Laplacian Eigenmaps is negligible.
-
-\paragraph{Testing.} 
-In our second experiment, we compare the Laplacian eigenmaps test $\varphi$ against the  population-level spectral projection test $\wt{\varphi}$.
-The setup, in terms of $n$ and $P$, is the same as that of our first experiment.
-To empirically evaluate the \textcolor{red}{critical radius} $\epsilon_n(\phi,H^1(\mc{X};M))$ of a test $\phi$, we compute $\phi$ for each $f_0 \in \mc{F} \subset H^1(\mc{X};M)$, where $\mc{F}$ is a discrete subset of $H^1(\mc{X};M)$.
-For each of $b = 1,2,\ldots,100$, we compute $\epsilon_b$, the smallest value of $\epsilon$ such that $R_n(\phi,\mc{F},\epsilon_b) \geq b/100$.
-Then we take $\bar{\epsilon} = 1/100 \cdot \sum_{b = 1}^{100}\epsilon_b$ to be our empirical measure of worst-case risk.
-In Figure~\ref{fig:fig2}, we see that the critical radii of both Laplacian eigenmaps and spectral projection tests are quite close to each other, and converge at roughly the minimax rate.
-\begin{figure*}[tb]
-	\includegraphics[width=.245\textwidth]{figures/testing/eigenfunction/critical_radius_by_sample_size_1d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/testing/eigenfunction/critical_radius_by_sample_size_1d_2s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/testing/eigenfunction/critical_radius_by_sample_size_2d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/testing/eigenfunction/critical_radius_by_sample_size_2d_2s.pdf}
-	\caption{Worst-case testing risk Laplacian eigenmaps (\texttt{LE}) and spectral projection (\texttt{SP}) tests, as a function of sample size $n$.
-Plots are on the same scale as Figure~\ref{fig:fig1}, and black line shows the minimax rate.
-All tests are set to have $.05$ Type I error, and are calibrated by simulation under the null.}
-	\label{fig:fig2}
-\end{figure*}
-
-\paragraph{Tuning parameters.}
-Our first two experiments demonstrate that in terms of statistical error, Laplacian eigenmaps methods are reasonable replacements for spectral projection methods.
-Laplacian eigenmaps depends on two tuning parameters, and in our final experiment we investigate the importance of both, focusing now on estimation.
-In Figure~\ref{fig:fig3}, we see how the mean-squared error of Laplacian eigenmaps changes as each tuning parameter is varied.
-As suggested by our theory, properly choosing the number of eigenvectors $K$ is crucial: the mean-squared error curves, as a function of $K$, always have a sharply defined minimum.
-On the other hand, as a function of the graph radius parameter $\varepsilon$ the mean-squared error curve is much closer to flat.
-This squares completely with our theory, which requires that the number of eigenvectors $K$ be much more carefully tuned that the graph radius $\varepsilon$.
-
-We also plot the out-of-sample mean squared error of Laplacian eigenmaps plus kernel smoothing, as a function of its various tuning parameters (which include the bandwidth $h$ as well as $\varepsilon$ and $K$.) Here the relationship between theory and empirics is more nuanced.
-On the one hand, empirically it seems that the optimal choice of bandwidth parameter $h$ is usually smaller than $\varepsilon$, as suggested by our theory.
-On the other hand, for Laplacian eigenmaps plus kernel smoothing we see that mean-squared error curves as a function of $K$ are often quite close to their minima even when we choose many more eigenvectors than is optimal for Laplacian eigenmaps or spectral projection.
-This is not reflected in our theory, where we require that $K$ be chosen in the same tight range as was required for Laplacian Eigenmaps to be optimal in-sample.
-However, it does make intuitive sense: extension by kernel smoothing further attenuates the noise, making the algorithm more forgiving to overfitting during the Laplacian Eigenmaps step.
-
-\begin{figure*}[tb]
-	\includegraphics[width=.245\textwidth]{figures/tuning/eigenfunction/mse_by_number_of_eigenvectors_1d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/tuning/eigenfunction/mse_by_radius_1d_1s.pdf} 
-	\includegraphics[width=.245\textwidth]{figures/tuning/eigenfunction/mse_by_number_of_eigenvectors_2d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/tuning/eigenfunction/mse_by_radius_2d_1s.pdf} 
-	\includegraphics[width=.245\textwidth]{figures/tuning/sobolev/mse_by_number_of_eigenvectors_1d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/tuning/sobolev/mse_by_radius_1d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/tuning/sobolev/mse_by_number_of_eigenvectors_2d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{figures/tuning/sobolev/mse_by_radius_2d_1s.pdf}  
-	\caption{Mean squared error of Laplacian Eigenmaps (\textcolor{red}{red}), Laplacian Eigenmaps plus kernel smoothing (\textcolor{blue}{blue}), and spectral projection (\textcolor{green}{green}) as a function of tuning parameters.
-Top row: the same regression function $f_0$ as used in Figure~\ref{fig:fig1}.
-Bottom row: the regression function $f_0 \propto \sum_{k} 1/\lambda_k^{1/2} \psi_k$.
-For all experiments, the sample size $n = 1000$, and the results are averaged over $200$ repetitions.
-In each panel, all tuning parameters except the one being varied are set to their optimal values.
-For Laplacian Eigenmaps plus kernel smoothing, circular points and a solid line are used to denote the error as a function of the graph radius $\varepsilon$, whereas triangular points and a dashed line are used to denote the error as a function of the bandwidth $h$.}
-	\label{fig:fig3}
-\end{figure*}
-
-\input{sections/discussion}
-
-% AG: I think this following might just be garbage, but I will leave it in in case it piques Ryan or Siva's interest.
-
-% Proposition~\ref{prop:graph_seminorm_ho}, and its proof, show that the iterated graph Laplacian $L_{n,\varepsilon}^s$ does a reasonable job of approximating a higher-order differential operator.
-The conditions we require --- that $\varepsilon$ be sufficiently large, that $f$ be zero-trace, and that $p$ have regularity of order $s - 1$ --- indicate that the iterated graph Laplacian is an imperfect, if adequate, tool for this job.
-This is no fault of the graph Laplacian.
-Rather, as pointed out by \textcolor{red}{(Sadhanala et al.
-2017)} its reflects that graph Laplacians were designed for a very general circumstance in which the only notions of derivation are first-differentials, divergences, and compositions of the two.
-Indeed, viewed in this light it is a pleasant surprise that iterated graph Laplacians approximate higher-order differentials at all.
-When one assumes Euclidean structure, as we do in this paper, there exist many methods (e.g.
-local linear embeddings, Hessian local linear embedding, local tangent space alignment) which leverage this structure to approximate differential operators in a more sophisticated manner.
-However, thus far little theoretical investigation has been done into even the pointwise behavior of these approaches.%
-
-\bibliographystyle{plainnat}
-\bibliography{../../graph_regression_bibliography}
-
+% Appendix
 \appendix
-
-\input{appendix.tex}
+\input{../appendix.tex}
 

### 2021-09-04 manuscripts/laplacian_eigenmaps/main_text.tex
Put Laplacian Eigenmaps paper into AOS format.

--- 2021-09-04 manuscripts/laplacian_eigenmaps/annals_of_statistics/laplacian_eigenmaps_regression_aos.tex
@@
-\documentclass[aos]{imsart}
-
-%% Packages
-\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
-% \RequirePackage[numbers]{natbib}
-\RequirePackage[authoryear]{natbib}
-\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
-\RequirePackage{graphicx}
-\usepackage{mathtools, enumitem}
-\usepackage{multirow}
-
-\startlocaldefs
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% For Axiom, Claim, Corollary, Hypothesis, %%
-%% Lemma, Theorem, Proposition              %%
-%% use \theoremstyle{plain}                 %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\theoremstyle{plain}
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{proposition}{Proposition}
-\theoremstyle{definition}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% For Assumption, Definition, Example,     %%
-%% Notation, Property, Remark, Fact         %%
-%% use \theoremstyle{remark}                %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\theoremstyle{remark}
-\newtheorem{remark}{Remark}
-\newtheorem{definition}{Definition}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Please put your definitions here:        %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%% Begin Ryan's definitions
-
-% Widebar
-\makeatletter
-\newcommand*\rel@kern[1]{\kern#1\dimexpr\macc@kerna}
-\newcommand*\widebar[1]{%
-	\begingroup
-	\def\mathaccent##1##2{%
-		\rel@kern{0.8}%
-		\overline{\rel@kern{-0.8}\macc@nucleus\rel@kern{0.2}}%
-		\rel@kern{-0.2}%
-	}%
-	\macc@depth\@ne
-	\let\math@bgroup\@empty \let\math@egroup\macc@set@skewchar
-	\mathsurround\z@ \frozen@everymath{\mathgroup\macc@group\relax}%
-	\macc@set@skewchar\relax
-	\let\mathaccentV\macc@nested@a
-	\macc@nested@a\relax111{#1}%
-	\endgroup
-}
-\makeatother
-
-% Min and max
-\newcommand{\argmin}{\mathop{\mathrm{argmin}}}
-\newcommand{\argmax}{\mathop{\mathrm{argmax}}}
-\newcommand{\minimize}{\mathop{\mathrm{minimize}}}
-\newcommand{\st}{\mathop{\mathrm{subject\,\,to}}}
-
-% Shortcuts
-\def\R{\mathbb{R}}
-
-%%% End Ryan's definitions
-
-%%% Begin Alden's additions
-\newcommand{\Ebb}{\mathbb{E}}
-\newcommand{\Pbb}{\mathbb{P}}
-\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
-\newcommand{\wt}[1]{\widetilde{#1}}
-\newcommand{\wh}[1]{\widehat{#1}}
-\newcommand{\mc}[1]{\mathcal{#1}}
-\newcommand{\Reals}{\mathbb{R}} % Same thing as Ryan's \R
-\newcommand{\Rd}{\Reals^d}
-\newcommand{\wb}[1]{\widebar{#1}}
-\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
-\newcommand{\Var}{\mathrm{Var}}
-\newcommand{\Cov}{\mathrm{Cov}}
-\newcommand{\1}{\mathbf{1}}
-\newcommand{\bj}{{\bf j}}
-\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
-
-\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
-\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
-\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
-\DeclareMathAccent{\wc}{0}{mathx}{"71}
-%%% End Alden's Additions
-
-\endlocaldefs
-
-\begin{document}
-
-\begin{frontmatter}
-
-\title{Minimax-optimal Regression over Sobolev Spaces via Laplacian Eigenmaps with Neighborhood Graphs}
-\runtitle{Minimax-optimal Laplacian Eigenmaps Regression}
-%\thankstext{T1}{A sample of additional note to the title.}
-
-\begin{aug}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Only one address is permitted per author. %%
-%% Only division, organization and e-mail is %%
-%% included in the address.                  %%
-%% Additional information can be included in %%
-%% the Acknowledgments section if necessary. %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author[A]{\fnms{Alden} \snm{Green}\ead[label=e1]{}},
-\author[A]{\fnms{Sivaraman} \snm{Balakrishnan}\ead[label=e1]{}}
-\and
-\author[A]{\fnms{Ryan J.} \snm{Tibshirani}\ead[label=e1]{}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Addresses                                %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\address[A]{Carnegie Mellon University, \printead{e1}}
-\end{aug}
-
-\begin{abstract}
-In this paper we study the statistical properties of nonparametric regression methods based on \emph{Laplacian eigenmaps}.
-These are spectral projection methods, which involve projecting a vector of observed responses onto a subspace spanned by certain eigenvectors of a neighborhood graph Laplacian.
-We show that in many cases these methods achieve minimax rates of convergence over continuum function classes.
-Specifically, we consider the problem of random design regression over Sobolev spaces $H^s(\mc{X})$---where $\mc{X} \subseteq \Rd$ is the support of the design density $p$---and show that under sufficient smoothness conditions on $p$, methods based on Laplacian eigenmaps achieve the optimal rates for both estimation (where the optimal rate is known to be $n^{-2s/(2s + d)}$) and goodness-of-fit testing ($n^{-4s/(4s + d)}$).
-We also show that Laplacian eigenmaps is \emph{manifold adaptive}: that is, we consider the situation where $\mc{X}$ is a manifold of small intrinsic dimension $m$, and give upper bounds establishing that Laplacian eigenmaps achieves the faster minimax estimation ($n^{-2s/(2s + m)}$) and testing ($n^{-4s/(4s + m)}$) rates of convergence.
-Finally, since the Laplacian eigenmaps estimator is defined only at observed design points, we propose an out-of-sample extension based on kernel smoothing, and prove that it achieves optimal out-of-sample error.
-We support these theoretical results with empirical evidence.
-\end{abstract}
-
-\begin{keyword}[class=MSC]
-\kwd{62G08, 62G10, 68R10}
-\end{keyword}
-
-\begin{keyword}
-\kwd{Laplacian eigenmaps, nonparametric regression, Sobolev space, minimax convergence rate}
-\end{keyword}
-
-\end{frontmatter}
-% Main text
 \section{Introduction}
@@
 More specifically, we consider the worst-case risk of Laplacian eigenmaps over normed balls in Sobolev spaces.
-Roughly speaking, for a domain $\mc{X} \subseteq \Rd$, the order-$s$ (Hilbert)-Sobolev space $H^s(\mc{X})$ consists of those functions which are $s$-times weakly differentiable, with order-$1$ up through order-$s$ derivatives bounded in $L^2(\mc{X})$ norm.
+Roughly speaking, for a domain $\mc{X} \subseteq \Rd$, the order-$s$ (Hilbert-)Sobolev space $H^s(\mc{X})$ consists of those functions which are $s$-times weakly differentiable, with order-$1$ up through order-$s$ derivatives bounded in $L^2(\mc{X})$ norm.
 Equivalently, Sobolev spaces can be defined in a spectral manner;  letting $\psi_1,\psi_2,\ldots$ be the eigenfunctions of a Laplacian operator defined with respect to $\mc{X}$, Sobolev spaces consist of those functions $f$ whose Fourier coefficients $\theta_k := \int_{\mc{X}} f \psi_k$ are bounded in a particular sequence norm (see Section~\ref{subsec:spectral_projection}).
@@
 
-To derive these results, we develop a framework for analysis reminiscent of that  classically used to analyze population-level spectral projection methods, but ultimately quite different.
+To derive these results, we develop a framework for analysis reminiscent of that  classically used to analyze population-level spectral projection estimators, but ultimately quite different.
 The chief difference is in the approximation error (bias).
@@
 Rates when $d > 4s$ assume that $f_0 \in L^4(\mc{X})$, and depend on $\|f_0\|_{L^4(\mc{X})}$.
-Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal.}
+Although we suppress it for simplicity, in all cases when othe Laplacian eigenmaps test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal.}
 	\label{tbl:testing_rates}
@@
 
-\subsection{Related work}
+\paragraph{Related work.}
 
@@
 
-\subsection{Roadmap}
+\paragraph{Roadmap.}
 We now outline the structure of the rest of this paper.
@@
 
-\subsection{Notation}
+\paragraph{Notation.}
 We frequently refer to various classical function classes.
@@
 
-\subsubsection{Flat Euclidean model}
+\paragraph{Flat Euclidean model.}
 In Definitions~\ref{def:model_flat_euclidean}-\ref{def:zero_trace_sobolev_space}, we collect the assumptions we make when working under the flat Euclidean model.
@@
 (Here $\partial/(\partial {\bf n})$ is the partial derivative operator in the direction of the normal vector $\mathbf{n}$.)
-
-\subsubsection{Manifold model}
+\paragraph{Manifold model.}
 As in the flat Euclidean case, we start with some regularity conditions on the design. One such condition will be on the reach $R$ of the manifold $\mc{X}$, which we recall is defined as follows:
@@
 
-\subsubsection{Estimation}
+\paragraph{Estimation.}
 We begin with an upper bound on the $L^2(P)$ risk of $\wt{f}$.
@@
 
-\begin{proof}{Proof of Proposition~\ref{prop:spectral_series_estimation}.}
+\paragraph{Proof of Proposition~\ref{prop:spectral_series_estimation}.}
 We decompose risk into squared bias and variance,
@@
 Otherwise, setting $K = \floor{M^2n}^{d/(2s + d)}$ balances squared bias and variance, and yields the claim.
-\end{proof}
-
-\subsubsection{Testing}
+\qed.
+
+\paragraph{Testing.}
 In the goodness-of-fit testing problem, one asks for a test function---formally, a Borel measurable function $\phi$ that takes values in $\{0,1\}$--- which can distinguish between the hypotheses
@@
 In other words, the regression testing problem over Sobolev spaces fundamentally changes when $s \leq d/4$, and hereafter when we discuss testing we will limit our consideration to $s > d/4$.
-\begin{proof}{Proof of Proposition~\ref{prop:spectral_series_testing}.}
+\paragraph{Proof of Proposition~\ref{prop:spectral_series_testing}.}
 We briefly lay out the main ideas needed to prove Proposition~\ref{prop:spectral_series_testing}, following the lead of~\cite{ingster2009} who prove a similar result in the special case where $M = 1$ and $P$ is the uniform distribution over $\mc{X} = [0,1]^d$, and referring to that work for more details.
@@
 Otherwise, choosing $K = \floor{M^2n}^{2d/(4s + d)}$ balances these two terms, and leads to~\eqref{eqn:spectral_series_testing}.
-\end{proof}
+\qed
 % (AG 8/19/21): Same comment as above, regarding the role of M in random design regression testing.
@@
 
-\subsubsection{Estimation} Laplacian eigenmaps depends on the kernel $\eta$ and two tuning parameters, the graph radius $\varepsilon$ and number of eigenvectors $K$.
+\paragraph{Estimation.} Laplacian eigenmaps depends on the kernel $\eta$ and two tuning parameters, the graph radius $\varepsilon$ and number of eigenvectors $K$.
 We will need to make some assumptions on each.
@@
 We comment on these assumptions after stating our first main theorem, regarding the estimation error of Laplacian eigenmaps.
-The proof of this theorem, along with the proofs of all subsequent results, can be found in the Supplementary Material.
+The proof of this theorem, along with the proofs of all subsequent results, can be found in the Appendix.
 \begin{theorem}
@@
 
-\subsubsection{Testing} Consider the test $\varphi = \1\{\wh{T} \geq t_{a}\}$, where $t_{a}$ is the threshold
+\paragraph{Testing.} Consider the test $\varphi = \1\{\wh{T} \geq t_{a}\}$, where $t_{a}$ is the threshold
 \begin{equation*}
@@
 
-\subsubsection{Estimation}
+\paragraph{Estimation.}
 In order to show that $\wh{f}$ is an optimal estimator over $H_0^s(\mc{X};M)$, we will require that $\varepsilon$ be meaningfully larger than the lower bound in~\ref{asmp:parameters_estimation_fo}.
@@
 As compared to a sparse $G$, this results in more non-zero entries in the graph Laplacian, and increases the computational burden involved in computing $\wh{f}$.
-To address this issue, in Appendix~H of the Supplementary Material we review some approaches to \emph{spectral sparsification}, in which one efficiently computes a sparse graph $\wc{G}$ that approximates $G$ in a spectral sense.
+To address this issue, in Appendix~\ref{sec:computational_considerations} we review some approaches to \emph{spectral sparsification}, in which one efficiently computes a sparse graph $\wc{G}$ that approximates $G$ in a spectral sense.
 The hope is that the Laplacian eigenmaps estimator $\wc{f}$ computed with respect to the sparsified graph $\wc{G}$ has similar statistical properties as $\wh{f}$, while being much faster to compute.
@@
 
-\subsubsection{Testing} The test $\varphi$ can adapt to the higher-order smoothness of $f_0$, when $\varepsilon$ and $K$ are chosen correctly.
+\paragraph{Testing.} The test $\varphi$ can adapt to the higher-order smoothness of $f_0$, when $\varepsilon$ and $K$ are chosen correctly.
 \begin{enumerate}[label=(P\arabic*)]
@@
 
-\subsubsection{Estimates of graph Sobolev seminorms}
+\paragraph{Estimates of graph Sobolev seminorms.}
 In Proposition~\ref{prop:graph_seminorm_fo} we restate an upper bound on the Dirichlet energy $\dotp{L_{n,\varepsilon}f}{f}_n$ from~\cite{green2021}. 
@@
 
-\subsubsection{Neighborhood graph eigenvalues}
+\paragraph{Neighborhood graph eigenvalues.}
 On the other hand, several recent works \citep{burago2014,garciatrillos18,calder2019} have analyzed the convergence of $\lambda_{k}$ towards $\lambda_{k}(\Delta_P)$.
@@
 
-\subsubsection{Empirical norm}
+\paragraph{Empirical norm.}
 Finally, in Proposition~\ref{prop:empirical_norm_sobolev} we establish that a one-sided bound of the form $\|f_0\|_n^2 \gtrsim \|f_0\|_P^2$ whenever $\|f_0\|_P^2$ is itself sufficiently large.
@@
 
-\subsubsection{Estimation}
+\paragraph{Estimation.}
 To ensure that $\wh{f}$ is an in-sample minimax rate-optimal estimator, we choose the kernel function $\eta$, graph radius $\varepsilon$ and number of eigenvectors $K$ as in~\ref{asmp:parameters_estimation_ho}, except with ambient dimension $d$ replaced by the intrinsic dimension $m$.
@@
 
-\subsubsection{Testing}
+\paragraph{Testing.}
 Likewise, to construct a minimax optimal test using $\wh{T}$, we choose $\varepsilon$ and $K$ as in~\ref{asmp:parameters_testing_fo}, except with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
@@
 
-\subsubsection{Analysis}
+\paragraph{Analysis.}
 The high-level strategy used to prove Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold} is the same as in the flat-Euclidean setting.
@@
 
-\subsection{Extension by kernel smoothing}
+\paragraph{Extension by kernel smoothing.}
 We now formally define our approach to extension by kernel smoothing.
@@
 
-\subsection{Out-of-sample error of kernel smoothed Laplacian eigenmaps}
+\paragraph{Out-of-sample error of kernel smoothed Laplacian eigenmaps.}
 
@@
 Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data, by a factor of $h^2$.
-On the other hand the bias term is typical, and when $\psi$ is an order-$s$ kernel a standard analysis (see Lemma~16 in the Supplementary Material) shows that $\|T_{P,h}f_0 - f_0\|_P^2 \lesssim |f|_{H^s(\mc{X})}^2 h^{2s}$.
+On the other hand the bias term is typical, and when $\psi$ is an order-$s$ kernel a standard analysis (see Lemma~\ref{lem:kernel_smoothing_bias}) shows that $\|T_{P,h}f_0 - f_0\|_P^2 \lesssim |f|_{H^s(\mc{X})}^2 h^{2s}$.
 \begin{enumerate}[label=(K\arabic*)]
@@
 
-\subsection{Estimation}
+\paragraph{Estimation.}
 In our first experiment, we compare the mean-squared error of the Laplacian eigenmaps estimator $\wh{f}$  to that of its population-level counterpart $\wt{f}$.
@@
 
-\subsection{Testing} 
+\paragraph{Testing.} 
 In our second experiment, we compare the Laplacian eigenmaps test $\varphi$ against the  population-level spectral projection test $\wt{\varphi}$.
@@
 
-\subsection{Tuning parameters}
+\paragraph{Tuning parameters.}
 Our first two experiments demonstrate that Laplacian eigenmaps methods have comparable statistical performance to spectral projection methods.
@@
 using the estimator $T_{n,h}{\bf Y}$ (possibly for a different choice of $h$).
-In Appendix~F.4 of the Supplementary Material, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{n,h}\wh{f}\|_P^2$ is of a strictly lower order than $\inf_{h'} \Ebb\|f_0 - T_{n,h'}{\bf Y}\|_P^2$.
+In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{n,h}\wh{f}\|_P^2$ is of a strictly lower order than $\inf_{h'} \Ebb\|f_0 - T_{n,h'}{\bf Y}\|_P^2$.
 This is possible because Laplacian eigenmaps induces a completely different bias than kernel smoothing.
@@
 Grounding each of these three points on a firmer and more complete theoretical basis would be, in our view, a valuable direction for future work.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Support information, if any,             %%
-%% should be provided in the                %%
-%% Acknowledgements section.                %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\begin{acks}[Acknowledgments]
-%
-%\end{acks}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Funding information, if any,             %%
-%% should be provided in the                %%
-%% funding section.                         %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\begin{funding}
-%\end{funding}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Supplementary Material, including data   %%
-%% sets and code, should be provided in     %%
-%% {supplement} environment with title      %%
-%% and short description. It cannot be      %%
-%% available exclusively as external link.  %%
-%% All Supplementary Material must be       %%
-%% available to the reader on Project       %%
-%% Euclid with the published article.       %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{supplement}
-\stitle{Supplement to ``Minimax-Optimal Regression Over Sobolev Spaces via Laplacian Eigenmaps with Neighborhood Graphs''.}
-\sdescription{The supplement contains all proofs omitted from the main text due to space constraints.}
-\end{supplement}
-
-% Bibliography
-\bibliographystyle{imsart-nameyear}
-\bibliography{../../../graph_regression_bibliography}
-
-\end{document}

### 2021-10-16 manuscripts/laplacian_eigenmaps/annals_of_statistics/laplacian_eigenmaps_regression_aos.tex
Updated Laplacian Eigenmaps manuscript.

@@
 
-\title{Minimax-optimal Regression over Sobolev Spaces via Laplacian Eigenmaps with Neighborhood Graphs}
-\runtitle{Minimax-optimal Laplacian Eigenmaps Regression}
+\title{Minimax optimal Regression over Sobolev Spaces via Laplacian Eigenmaps with Neighborhood Graphs}
+\runtitle{Minimax optimal Laplacian Eigenmaps Regression}
 %\thankstext{T1}{A sample of additional note to the title.}
@@
 The chief difference is in the approximation error (bias).
-For our purposes we must establish that with high probability, the eigenvectors of graph Laplacians (which are data-dependent and hence random objects) can be used to effectively approximate continuum Sobolev functions, whereas in the classical analysis an upper bound on the approximation error follows immediately from the aforementioned spectral characterization of Sobolev spaces.
+For our purposes we must establish that with high probability, the eigenvectors of graph Laplacians (which are data-dependent and hence random objects) can be used to effectively approximate continuum Sobolev functions.
+In contrast, for the classical analysis of population-level spectral projection methods, an upper bound on the approximation error follows immediately from the aforementioned spectral characterization of Sobolev spaces.
 To tightly upper bound the approximation error of Laplacian eigenmaps, we rely on several recent results regarding concentration of the spectra of graph Laplacians around their population limits, and prove some additional results of this type which may be of independent interest.
@@
 It is equal to $d$ under the assumptions of Model~\ref{def:model_flat_euclidean}, and equal to $m$ under the assumptions of Model~\ref{def:model_manifold}.
-The pre-factor $(n\varepsilon^{2 + \mathrm{dim}(\mc{X})})^{-1}$ ensures non-degenerate stable limits as $n \to \infty, \varepsilon \to 0$).
+The pre-factor $(n\varepsilon^{2 + \mathrm{dim}(\mc{X})})^{-1}$ ensures non-degenerate limits as $n \to \infty, \varepsilon \to 0$).
 Note that $(n\varepsilon^{\dim(\mc{X}) + 2}) \cdot L_{n,\varepsilon} = D - W$, where $D \in \Reals^{n \times n}$ is the diagonal degree matrix, $D_{ii} = \sum_{i = 1}^{n} W_{ij}$.
@@
 
-\subsection{Spectral Projection using Continuum Eigenfunctions}
+\subsection{Spectral projection using continuum eigenfunctions}
 \label{subsec:spectral_projection}
@@
 
-\begin{proof}{Proof of Proposition~\ref{prop:spectral_series_estimation}.}
+\begin{proof}[Proof of Proposition~\ref{prop:spectral_series_estimation}.]
 We decompose risk into squared bias and variance,
@@
 In other words, the regression testing problem over Sobolev spaces fundamentally changes when $s \leq d/4$, and hereafter when we discuss testing we will limit our consideration to $s > d/4$.
-\begin{proof}{Proof of Proposition~\ref{prop:spectral_series_testing}.}
+\begin{proof}[Proof of Proposition~\ref{prop:spectral_series_testing}.]
 We briefly lay out the main ideas needed to prove Proposition~\ref{prop:spectral_series_testing}, following the lead of~\cite{ingster2009} who prove a similar result in the special case where $M = 1$ and $P$ is the uniform distribution over $\mc{X} = [0,1]^d$, and referring to that work for more details.
@@
 
-The main takeaway from Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} is that spectral projection methods for nonparametric regression achieve optimal rates of convergence, when the regression function $f_0$ is Sobolev smooth and the design density $p$ is known a priori and satisfies an appropriate notion of smoothness\footnote{The assumption $p \in C^{\infty}(\mc{X})$ could likely be weakened, but since this would not substantially add to the main points of Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, we do not pursue the details further.}.
-As we will see, Laplacian eigenmaps achieves comparable rates of convergence when $p$ is sufficiently smooth but potentially unknown, and in this sense both learns and leverages the design distribution in a way that more classical spectral projection methods do not.
+The main takeaway from Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} is that spectral projection methods for nonparametric regression achieve optimal rates of convergence, when the regression function $f_0$ is Sobolev smooth and the design density $p$ is known a priori and satisfies an appropriate notion of smoothness.\footnote{The assumption $p \in C^{\infty}(\mc{X})$ could likely be weakened, but since this would not substantially add to the main points of Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, we do not pursue the details further.} As we will see, Laplacian eigenmaps achieves comparable rates of convergence when $p$ is sufficiently smooth but potentially unknown, and in this sense both learns and leverages the design distribution in a way that more classical spectral projection methods do not.
 
@@
 
-\section{Minimax Optimality of Laplacian Eigenmaps}
+\section{Minimax optimality of Laplacian eigenmaps}
 \label{sec:minimax_optimal_laplacian_eigenmaps}
@@
 We return to the question of out-of-sample estimation later in Section~\ref{sec:out_of_sample}.
-
-	There is one subtlety introduced by the use of in-sample mean squared error.
+However, there is one subtlety introduced by the use of in-sample mean squared error.
 Since elements $f \in H^s(\mc{X})$ are equivalence classes, defined only up to a set of measure zero, one cannot really speak of the pointwise evaluation $f_0(X_i)$, as we do by defining our target of estimation to be $(f_0(X_1),\ldots,f_0(X_n))$, until one selects a representative of each equivalence class $f$.
-Implicitly, we will always pick the \emph{precise representative} $f_0^{\ast} \in f_0$ (as defined in~\cite{evans15}), and the notation ``$f_0(X_i)$'' should always be interpreted as $f_0^{\ast}(X_i)$.
+Implicitly, we will always pick the \emph{precise representative} $f_0^{\ast} \in f_0$ (as defined in~\cite{evans15}), and the notation ``$f_0(X_i)$'' should always be interpreted as the evaluation $f_0^{\ast}(X_i)$.
 To be clear, however, it does not really matter which representative we choose, since all versions agree except on a set of measure zero, and so any two $g_0,h_0 \in f_0$ satisfy $g_0(X_i) = h_0(X_i)$ for all $i = 1,\ldots,n$ almost surely.
@@
 It is intuitively reasonable that data-dependent spectral projection methods, such as Laplacian eigenmaps, should have more difficulty achieving the minimax rates of convergence for testing, as opposed to estimation.
-Optimally balancing the bias and variance of spectral projection tests (for instance~\eqref{pf:spectral_series_test}) requires choosing more and higher-frequency eigenvectors than in estimation.
-It is known~\citep{trillos2019} that higher-frequency graph Laplacian eigenvectors---which correspond to larger eigenvalues---are meaningfully worse approximations of their population-level limits.
+Optimally balancing the bias and variance of spectral projection tests requires choosing more and higher-frequency eigenvectors than in estimation.
+It is known~\citep{trillos2019} that at a given sample size $n$, higher-frequency graph Laplacian eigenvectors---which correspond to larger eigenvalues---are meaningfully worse approximations of their population-level limits.
 Put more simply, minimax optimal testing when $d > 4$ appears to require stronger approximation properties than graph Laplacian eigenvectors can satisfy.
@@
 	\item Finally, we combine the results of the previous two steps to deduce an upper bound on $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$ in terms of the squared Sobolev norm $\|f\|_{H^s(\mc{X})}^2$.
-Roughly speaking, when $s$ is odd, $\dotp{L_{P,\varepsilon}^sf}{f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}^jf) \approx \sigma_{\eta}^{2j}E_{P,\varepsilon}(\Delta_P^jf)$, whereas when $s$ is even $\dotp{L_{P,\varepsilon}^sf}{f}_P = \|L_{P,\varepsilon}L_{P,\varepsilon}^{j}f\|_{P}^2 \approx \sigma_{\eta}^{2j}\|L_{P,\varepsilon} \Delta_Pf\|_P^2$.
-Reasoning in this way, we can translate estimates of $L_{P,\varepsilon}^jf$ into an upper bound on the order-$s$ non-local Sobolev seminorm, even though $s > j$.
+The nature of this last step  depends on whether $s$ is an even or odd integer.
+Roughly speaking, when $s$ is odd, letting $j = (s - 1)/2$, we show that
+	$$\dotp{L_{P,\varepsilon}^sf}{f}_P = \dotp{L_{P,\varepsilon}L_{P,\varepsilon}^jf}{L_{P,\varepsilon}^jf}_P \approx \sigma_{\eta}^{2j}\dotp{L_{P,\varepsilon}\Delta_P^jf}{\Delta_P^j}_{P} \lesssim \sigma_{\eta}^{2j + 1} \dotp{\Delta_P^jf}{f}_{P}.$$
+	When $s$ is even, letting $j = (s - 2)/2$, we show that
+	\begin{equation*}
+	\dotp{L_{P,\varepsilon}^sf}{f}_P = \|L_{P,\varepsilon}L_{P,\varepsilon}^{j}f\|_{P}^2 \approx \sigma_{\eta}^{2j}\|L_{P,\varepsilon} \Delta_P^jf\|_P^2 \lesssim \|\Delta_P^{j + 1}f\|_P^2.
+	\end{equation*}
+	The desired upper bound then follows from the boundedness of the density $p$.
 \end{enumerate}
-Together, these steps establish Proposition~\ref{prop:graph_seminorm_ho}.
-It is worth pointing out that we do not try to establish the pointwise estimate $L_{P,\varepsilon}^sf \to \sigma_{\eta}^s\Delta_{P}^sf$ in $L^2(P)$.
+It is worth pointing out that we do not need to establish the pointwise estimate $L_{P,\varepsilon}^sf \to \sigma_{\eta}^s\Delta_{P}^sf$ in $L^2(P)$.
 If we had such an estimate, it would immediately follow that $\dotp{L_{P,\varepsilon}^sf}{f}_{P} \to \sigma_{\eta}^s\dotp{\Delta_P^sf}{f}_{P}$.
@@
 Combining these statements with~\eqref{eqn:graph_eigenvalue}, we conclude that $\lambda_{K} = \Omega_P(K^{2/d})$ so long as $K \lesssim \varepsilon^{-d}$.
+We note that Lemma~2 of~\cite{green2021} actually gives a corresponding upper bound on the graph eigenvalues $\lambda_k$, but we will not need it.
 
@@
 
-\section{Manifold Adaptivity}
+\section{Manifold adaptivity}
 \label{sec:manifold_adaptivity}
@@
 
-As discussed previously, when $\mc{X}$ is a domain without boundary and $\Delta_P$ is the manifold weighted Laplace-Beltrami operator, appropriate bounds on the graph eigenvalues $\lambda_k$ have already been derived in \citep{burago2014,trillos2019,garciatrillos19}.
-The precise result we need is a direct consequence of Theorem 2.4 of~\citep{calder2019}.
-\begin{proposition}[\textbf{c.f Theorem 2.4 of~\citep{calder2019}}]
+As discussed previously, when $\mc{X}$ is a domain without boundary and $\Delta_P$ is the manifold weighted Laplace-Beltrami operator, appropriate bounds on the graph eigenvalues $\lambda_k$ have already been derived in \citet{burago2014,trillos2019,garciatrillos19}.
+The precise result we need is a direct consequence of Theorem 2.4 of~\citet{calder2019}.
+\begin{proposition}[{cf. Theorem 2.4 of~\citet{calder2019}}]
 	\label{prop:graph_eigenvalue_manifold}
@@
 	\item \emph{Alternatives to extension by kernel smoothing}.
-There exist other approaches to extending a function $f$ from its evaluations $\{f(X_1),\ldots,f(X_n)\}$: for instance, minimum-norm interpolation in some RKHS or Banach space or 1-nearest neighbors regression.
+There exist other approaches to extending a function $f$ from its evaluations $(f(X_1),\ldots,f(X_n))$: for instance, minimum-norm interpolation in some RKHS or Banach space or 1-nearest neighbors regression.
 We consider extension by kernel smoothing because it is a simple and statistically optimal procedure that does not require any knowledge of the domain $\mc{X}$ or distribution $P$---as we have argued, this latter property is one of the main selling points of Laplacian eigenmaps as a tool for nonparametric regression.
@@
 We see that all estimators have mean-squared error converging to zero at roughly the minimax rate.
-While unsurprisingly population-level spectral projection has smaller error, generally speaking the error of Laplacian eigenmaps approaches that of population-level spectral projection as $n$ gets larger.
+While unsurprisingly population-level spectral projection has smaller error than Laplacian eigenmaps, generally speaking the error of eigenmaps approaches that of population-level spectral projection as $n$ gets larger.
 
@@
 First we take $\mc{F} = \{M/\lambda_k^{s/2} \psi_k\}_{k = 1}^{n}$ to be a discrete subset of $H^1(\mc{X};M)$.
-Then, for each $f_0 \in \mc{F}$, we run a given test $\phi$ (either the Laplacian eigenmaps test $\phi = \varphi$, or the population-level spectral projection test $\phi = \wt{\varphi}$.) and record whether it was a false negative or true positive.
+Then, for each $f_0 \in \mc{F}$, we run a given test $\phi$ (either the Laplacian eigenmaps test $\phi = \varphi$ or the population-level spectral projection test $\phi = \wt{\varphi}$) and record whether it was a false negative or true positive.
 We repeat this process over $100$ replications, giving a Monte Carlo estimate of the type II error $E_{f_0}[1 - \phi]$ for each $f_0 \in \mc{F}$.

### 2021-10-16 manuscripts/laplacian_eigenmaps/annals_of_statistics/supplement.tex
Updated Laplacian Eigenmaps manuscript.

@@
 
-\title{Supplementary Material For: Minimax-optimal Regression over Sobolev Spaces via Laplacian Eigenmaps with Neighborhood Graphs}
+\title{Supplementary Material For: Minimax optimal Regression over Sobolev Spaces via Laplacian Eigenmaps with Neighborhood Graphs}
 %\thankstext{T1}{A sample of additional note to the title.}
@@
 In the rest of this section, we will mildly abuse notation and write $f_0 = (f_0(x_1),\ldots,f_0(x_n)) \in \Reals^n$.
-We will also write ${\bf Y} = (Y_1,\ldots,Y_n)$.
+We will also write ${\bf Y} = (Y_1,\ldots,Y_n)$, and ${\bf w} = (w_1,\ldots,w_n)$.
 
@@
 \end{lemma}
-\paragraph{Proof (of Lemma~\ref{lem:fixed_graph_estimation}).}
+\paragraph{Proof (of Lemma~\ref{lem:fixed_graph_estimation})}
 By the triangle inequality,
@@
 Otherwise the second   in~\eqref{pf:fixed_graph_estimation_1} (estimation error) is random.
-Observe that $\dotp{v_k}{\varepsilon}_n \overset{d}{=} Z_k/\sqrt{n}$, where $(Z_1,\ldots,Z_n) \sim N(0,I_{n \times n})$.
+Observe that $\dotp{v_k}{{\bf w}}_n \overset{d}{=} Z_k/\sqrt{n}$, where $(Z_1,\ldots,Z_n) \sim N(0,I_{n \times n})$.
 Again using the orthonormality of the eigenvectors $v_k$, we have
 \begin{equation*}
-\|\wh{f} - \mathbb{E}\wh{f}\|_n^2 = \sum_{k = 1}^{K} \dotp{v_k}{\varepsilon}_n^2 \overset{d}{=} \frac{1}{n}\sum_{k = 1}^{K} Z_k^2.
+\|\wh{f} - \mathbb{E}\wh{f}\|_n^2 = \sum_{k = 1}^{K} \dotp{v_k}{{\bf w}}_n^2 \overset{d}{=} \frac{1}{n}\sum_{k = 1}^{K} Z_k^2.
 \end{equation*}
@@
 \end{lemma}
-\paragraph{Proof (of Lemma~\ref{lem:fixed_graph_testing}).}
+\paragraph{Proof (of Lemma~\ref{lem:fixed_graph_testing})}
 We first compute the expectation and variance of $\wh{T}$, then apply Chebyshev's inequality to upper bound the Type I and Type II error.
@@
 \begin{equation*}
-\Ebb[\wh{T}] = \sum_{k = 1}^{K} \Ebb[\dotp{Y}{v_k}_n^2] = \sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 + \Ebb[2\dotp{f_0}{v_k}_n\dotp{\varepsilon}{v_k}_n + \dotp{\varepsilon}{v_k}_n^2] = \frac{K}{n} + \sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2.
+\Ebb[\wh{T}] = \sum_{k = 1}^{K} \Ebb[\dotp{Y}{v_k}_n^2] = \sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 + \Ebb[2\dotp{f_0}{v_k}_n\dotp{{\bf w}}{v_k}_n + \dotp{{\bf w}}{v_k}_n^2] = \frac{K}{n} + \sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2.
 \end{equation*}
@@
 \underline{\emph{Variance}.}
-Recall from the proof of Lemma~\ref{lem:fixed_graph_estimation} that $\dotp{\varepsilon}{v_k}_n \overset{d}{=} Z_k/\sqrt{n}$ for $(Z_1,\ldots,Z_n) \sim N(0,I_{n \times n})$.
+Recall from the proof of Lemma~\ref{lem:fixed_graph_estimation} that $\dotp{{\bf w}}{v_k}_n \overset{d}{=} Z_k/\sqrt{n}$ for $(Z_1,\ldots,Z_n) \sim N(0,I_{n \times n})$.
 Expanding the square, and recalling that $\Cov[Z,Z^2] = 0$ for Gaussian random variables, we have that
@@
 
-\paragraph{Proof (of Proposition~4).}
+\paragraph{Proof (of Proposition~4)}
 Proposition~4 follows immediately from~\eqref{pf:graph_seminorm_ho_1} and~\eqref{pf:graph_seminorm_ho_2}. \qed
@@
 Then~\ref{pf:graph_seminorm_ho_1} follows immediately from Lemma~\ref{lem:graph_seminorm_bias}, by Markov's inequality.
-\paragraph{Proof (of Lemma~\ref{lem:graph_seminorm_bias}).}
+\paragraph{Proof (of Lemma~\ref{lem:graph_seminorm_bias})}
 We begin by introducing some notation.
@@
 
-\paragraph{Proof (of Lemma~\ref{lem:approximation_error_nonlocal_laplacian}).}
+\paragraph{Proof (of Lemma~\ref{lem:approximation_error_nonlocal_laplacian})}
 Throughout this proof, we shall assume that $f$ and $p$ are smooth functions, meaning they belong to $C^{\infty}(\mc{X})$.
@@
 g_{1,1}(x) + \frac{1}{2}g_{2,0}(x) & = \int_{\mc{X}} d_x^{1}f(z) d_x^{1}p(z) \eta(\|z\|) \,dz + \frac{p(x)}{2} \int_{\mc{X}} d_x^{2}f(z) \eta(\|z\|) \,dz \\
-& = \sum_{i_1 = 1}^{d} \sum_{i_2 = 1}^{d} D^{e_{i_1}}f(x) D^{e_{i_2}}p(x) \int_{\mc{X}} z^{e_{i_1} + e_{i_2}} \eta(\|z\|) \,dz + \frac{p(x)}{2} \sum_{i_1 = 1}^{d} \sum_{i_2 = 1}^{d}D^{e_{i_2}+e_{i_2}}f(x)\int_{\mc{X}} z^{e_{i_1} + e_{i_2}} \eta(\|z\|) \,dz\\
+& = \sum_{i_1 = 1}^{d} \sum_{i_2 = 1}^{d} D^{e_{i_1}}f(x) D^{e_{i_2}}p(x) \int_{\mc{X}} z^{e_{i_1} + e_{i_2}} \eta(\|z\|) \,dz \\
+& \quad + \frac{p(x)}{2} \sum_{i_1 = 1}^{d} \sum_{i_2 = 1}^{d}D^{e_{i_2}+e_{i_2}}f(x)\int_{\mc{X}} z^{e_{i_1} + e_{i_2}} \eta(\|z\|) \,dz\\
 & = \sum_{i = 1}^{d} D^{e_{i}}f(x) D^{e_{i}}p(x) \int_{\mc{X}} z^2 \eta(\|z\|) \,dz + \frac{p(x)}{2} \sum_{i = 1}^{d} D^{2e_{i}}f(x)\int_{\mc{X}} z^2 \eta(\|z\|) \,dz\\ 
@@
 \end{equation}
-\paragraph{Case 1: $s$ odd.}
+\paragraph{Case 1: $s$ odd}
 Suppose $s$ is odd, so that $s \geq 3$.
@@
 \begin{align*}
-\frac{2}{\varepsilon^{d + 2}}\int_{\mc{X}_{k\varepsilon}} \int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_P^kf(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) & \leq \frac{2p_{\max}}{\varepsilon^{2}}\int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_P^kf(x)\bigr)^2 \,dP(x) \\
-& \leq C_1M^2.
+& \frac{2}{\varepsilon^{d + 2}}\int_{\mc{X}_{k\varepsilon}} \int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_P^kf(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) \\
+& \quad \leq \frac{2p_{\max}}{\varepsilon^{2}}\int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_P^kf(x)\bigr)^2 \,dP(x) \\
+& \quad \leq C_1M^2.
 \end{align*}
@@
 \end{equation*}
-\paragraph{Case 2: $s$ even.}
+\paragraph{Case 2: $s$ even}
 If $s \in \mathbb{N}$ is even, $s \geq 2$, then letting $k = (s - 2)/2$, the  self-adjointness of $L_{P,\varepsilon}$ implies
@@
 
-\paragraph{Proof (of Lemma~\ref{lem:l2estimate_nonlocal_laplacian}).}
+\paragraph{Proof (of Lemma~\ref{lem:l2estimate_nonlocal_laplacian})}
 We fix a version of $f \in L^2(U)$, so that we may speak of its pointwise values.
@@
 \begin{equation}
-\int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx = 2 \int_{U} \int_{U} |f(x)|^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx \leq 2 \varepsilon^d \int_{U} |f(x)|^2 \,dx = 2\varepsilon^d \|f\|_{L^2(U)}^2,
+\begin{aligned}
+\int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx & = 2 \int_{U} \int_{U} |f(x)|^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx \\
+& \leq 2 \varepsilon^d \int_{U} |f(x)|^2 \,dx = 2\varepsilon^d \|f\|_{L^2(U)}^2,
 \label{pf:l2estimate_nonlocal_laplacian_2}
+\end{aligned}
 \end{equation}
@@
 
-\paragraph{Proof (of Lemma~\ref{lem:dirichlet_estimate_nonlocal_laplacian}).}
+\paragraph{Proof (of Lemma~\ref{lem:dirichlet_estimate_nonlocal_laplacian})}
 We have
 \begin{equation*}
-E_{P,\varepsilon}(f) = \frac{1}{\varepsilon^{2 + d}} \int_{U} \int_{U} \bigl(f(x) - f(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) \leq \frac{2p_{\max}^2}{\varepsilon^{2 + d}} \int_{U} \int_{U} \bigl(|f(x)|^2 + |f(x')|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx,
+\begin{aligned}
+E_{P,\varepsilon}(f) & = \frac{1}{\varepsilon^{2 + d}} \int_{U} \int_{U} \bigl(f(x) - f(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) \\
+& \leq \frac{2p_{\max}^2}{\varepsilon^{2 + d}} \int_{U} \int_{U} \bigl(|f(x)|^2 + |f(x')|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx,
+\end{aligned}
 \end{equation*}
@@
 
-\paragraph{Proof (of Lemma~\ref{lem:graph_seminorm_bias2}).}
+\paragraph{Proof (of Lemma~\ref{lem:graph_seminorm_bias2})}
 Let $G_{n,\varepsilon}[X_{i\bj}]$ be the subgraph induced by vertices $X_i, X_{\bj_1},\ldots,X_{\bj_s}$.
@@
 \begin{align}
-\mathbb{E}\Bigl[ |f(X_j) - f(X_{i})|^2 \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr] & \leq \bigl((s + 1)p_{\max}\nu_d\varepsilon^d\bigr)^{|i\bj\setminus\{j \cup i\}|} \cdot \mathbb{E}\Bigl[ |f(X_j) - f(X_{i})|^2 \1\{\|X_i - X_j\| \leq \varepsilon\}\Bigr] \nonumber \\
-& \leq \bigl((s + 1)p_{\max}\nu_d\varepsilon^d\bigr)^{|i\bj\setminus\{j \cup i\}|} \cdot p_{\max}^2 \nu_d \varepsilon^{2 + d} M^2 \label{pf:graph_seminorm_bias2_1}
+& \mathbb{E}\Bigl[ |f(X_j) - f(X_{i})|^2 \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr] \nonumber\\
+& \quad \leq \bigl((s + 1)p_{\max}\nu_d\varepsilon^d\bigr)^{|i\bj\setminus\{j \cup i\}|} \cdot \mathbb{E}\Bigl[ |f(X_j) - f(X_{i})|^2 \1\{\|X_i - X_j\| \leq \varepsilon\}\Bigr] \nonumber \\
+& \quad \leq \bigl((s + 1)p_{\max}\nu_d\varepsilon^d\bigr)^{|i\bj\setminus\{j \cup i\}|} \cdot p_{\max}^2 \nu_d \varepsilon^{2 + d} M^2 \label{pf:graph_seminorm_bias2_1}
 \end{align}
@@
 
-\paragraph{Proof (of Proposition~7).} Follows immediately from~\eqref{pf:graph_seminorm_manifold_1} and~\eqref{eqn:manifold_nonlocal_seminorm}. \qed
+\paragraph{Proof (of Proposition~7)} Follows immediately from~\eqref{pf:graph_seminorm_manifold_1} and~\eqref{eqn:manifold_nonlocal_seminorm}. \qed
 
@@
 From here forward we will assume $\varepsilon < R/2$. 
-\paragraph{Upper bound on $A_1(x)$.}
+\paragraph{Upper bound on $A_1(x)$}
 Consequently $\eta(\|x' - x\|/\varepsilon) \geq \eta(d_{\mc{X}}(x',x)/\varepsilon)$.
@@
 \begin{equation*}
-A_1(x) \leq \frac{8L_{\eta}\varepsilon^2}{R^2}\int_{\mc{X}}\bigl[f(x') - f(x)\bigr]^2 \1\{\|x' - x\| \leq \varepsilon\} \,d\mu(x') + \|\eta\|_{\infty} \int_{\mc{X}}\bigl[f(x') - f(x)\bigr]^2 \1\bigl\{\varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon}\bigr\} \,d\mu(x') \\
+\begin{aligned}
+A_1(x) & \leq \frac{8L_{\eta}\varepsilon^2}{R^2}\int_{\mc{X}}\bigl[f(x') - f(x)\bigr]^2 \1\{\|x' - x\| \leq \varepsilon\} \,d\mu(x') \\
+& \quad + \|\eta\|_{\infty} \int_{\mc{X}}\bigl[f(x') - f(x)\bigr]^2 \1\bigl\{\varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon}\bigr\} \,d\mu(x')
+\end{aligned}
 \end{equation*}
@@
 
-\paragraph{Upper bound on $A_2(x)$.}
+\paragraph{Upper bound on $A_2(x)$}
 Integrating over $x' \in \mc{X}$, we see that
 \begin{align}
-\int_{\mc{X}} \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') & \leq \frac{8L_{\eta}\varepsilon^2}{R^2} \int_{\mc{X}} \1\bigl\{d_{\mc{X}}(x',x)\bigr\} \,d\mu(x') + p_{\max}\|\eta\|_{\infty} \int_{\mc{X}} \1\bigl\{ \varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon} \bigr\} \,d\mu(x') \nonumber \\
+\int_{\mc{X}} \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') & \leq \frac{8L_{\eta}\varepsilon^2}{R^2} \int_{\mc{X}} \1\bigl\{d_{\mc{X}}(x',x)\bigr\} \,d\mu(x')\nonumber \\
+& \quad + p_{\max}\|\eta\|_{\infty} \int_{\mc{X}} \1\bigl\{ \varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon} \bigr\} \,d\mu(x') \nonumber \\
 & = \frac{8L_{\eta}\varepsilon^2}{R^2} \cdot \mu\bigl(B(x,\varepsilon)\bigr) +  p_{\max}\|\eta\|_{\infty} \Bigl[\mu\bigl(B(x,\wt{\varepsilon})\bigr) - \mu\bigl(B(x,\varepsilon)\bigr) \Bigr].
@@
 \begin{align*}
-\int_{\mc{X}} \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') & \leq  \frac{8L_{\eta}\varepsilon^2}{R^2}\Bigl[\omega_m\varepsilon^{m} + CmK_0\varepsilon^{m + 2}\Bigr] + \|\eta\|_{\infty} \Bigl[\omega_m(\wt{\varepsilon}^m - {\varepsilon}^m) + 2CmK_0\varepsilon^{m + 2}\Bigr] \\
+\int_{\mc{X}} \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') & \leq  \frac{8L_{\eta}\varepsilon^2}{R^2}\Bigl[\omega_m\varepsilon^{m} + CmK_0\varepsilon^{m + 2}\Bigr] \nonumber \\
+& \quad + \|\eta\|_{\infty} \Bigl[\omega_m(\wt{\varepsilon}^m - {\varepsilon}^m) + 2CmK_0\varepsilon^{m + 2}\Bigr] \\
 & \leq \frac{8L_{\eta}\varepsilon^2}{R^2}\Bigl[\omega_m\varepsilon^{m} + R^2CmK_0\varepsilon^{m}\Bigr] + \|\eta\|_{\infty} \varepsilon^{m + 2}\Bigl[\frac{27\omega_m}{R^2} + 2CmK_0\Bigr] \\
@@
 
-\paragraph{Putting together the pieces.}
+\paragraph{Putting together the pieces}
 Plugging our upper bounds on $A_1(x)$ and $A_2(x)$ back into~\eqref{pf:error_euclidean_distance_0}, we deduce that
@@
 
-\paragraph{Case 1: $s = 2$.}
+\paragraph{Case 1: $s = 2$}
 When $s = 2$, the triangle inequality implies that
@@
 
-\paragraph{Case 2: $s = 3$.}
+\paragraph{Case 2: $s = 3$}
 When $s = 3$, the triangle inequality implies that
@@
 
-\paragraph{Proof (of Lemmas~\ref{lem:dirichlet_energy_l2} and~\ref{lem:dirichlet_energy_sobolev}).}
+\paragraph{Proof (of Lemmas~\ref{lem:dirichlet_energy_l2} and~\ref{lem:dirichlet_energy_sobolev})}
 Define the non-local energy $\wt{E}_{P,\varepsilon}$ with respect to geodesic distance,
@@
 
-\paragraph{Proof (of Lemma~\ref{lem:dirichlet_energy_remainder}).}
+\paragraph{Proof (of Lemma~\ref{lem:dirichlet_energy_remainder})}
 Following exactly the steps of the proof of Lemma~3.3 of \citet{burago2014}, but replacing all references to a ball of radius $r$ by references to the set difference between balls of radius $\wt{\varepsilon}$ and $\varepsilon$, we obtain that
@@
 
-\paragraph{Proof (of Lemma~\ref{lem:manifold_graph_seminorm_bias2}).}
+\paragraph{Proof (of Lemma~\ref{lem:manifold_graph_seminorm_bias2})}
 The proof of Lemma~\ref{lem:manifold_graph_seminorm_bias2} is identical to the proof of Lemma~\ref{lem:graph_seminorm_bias2}, upon substituting the ambient dimension $m$ for the intrinsic dimension $d$, and using Lemma~\ref{lem:dirichlet_energy_sobolev} rather than Lemma~\ref{lem:dirichlet_estimate_nonlocal_laplacian} to establish~\eqref{pf:graph_seminorm_bias2_1}.
@@
 
-\paragraph{Proof (of Proposition~6).}
+\paragraph{Proof (of Proposition~6)}
 Rearranging~\eqref{eqn:gagliardo_nirenberg} and raising both sides to the $4$th power, we see that
@@
 
-\paragraph{Proof of Theorem~1.}
+\paragraph{Proof of Theorem~1}
 We condition on the event that the design points $X_1,\ldots,X_n$ satisfy
@@
 
-\paragraph{Proof of Theorem~3.}
+\paragraph{Proof of Theorem~3}
 Follows identically to the proof of Theorem~1, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~4 rather than Proposition~3 and Assumption~(P3) rather than Assumption~(P1).
 
-\paragraph{Proof of Theorem~6.}
+\paragraph{Proof of Theorem~6}
 Follows identically to the proof of Theorem~1, substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~7 rather than Proposition~3, Proposition~8 rather than Proposition~4, and Assumption~(P6) rather than Assumption~(P1).
@@
 
-\paragraph{Proof of Theorem~2.}
+\paragraph{Proof of Theorem~2}
 We have already upper bounded the Type I error of $\varphi$ in Lemma~\ref{lem:fixed_graph_testing}, and it remains to upper bound the Type II error.
@@
 
-\paragraph{Proof of Theorem~4.}
+\paragraph{Proof of Theorem~4}
 Follows identically to the proof of Theorem~1, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~4 rather than Proposition~3 and Assumption~(P4) rather than Assumption~(P2).
 
-\paragraph{Proof of Theorem~7.}
+\paragraph{Proof of Theorem~7}
 Follows identically to the proof of Theorem~1, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~7 rather than Proposition~3, Proposition~8 rather than Proposition~4, Proposition~9 rather than Proposition~6, and Assumption~(P6) rather than Assumption~(P2).
 
-\paragraph{Proof of Theorem~5.}
+\paragraph{Proof of Theorem~5}
 Note that our choices of $K$ and $\varepsilon$ ensure  that~\eqref{pf:laplacian_eigenmaps_testing_fo_1} (with $L_{n,\varepsilon}^s$ replacing $L_{n,\varepsilon}$) and~\eqref{pf:laplacian_eigenmaps_testing_fo_2} are satisfied with probability at least $1 - b/2$.
@@
 \Ebb\biggl[\int \Bigl(\wt{L}_{n,h}f_0(x)\Bigr)^2 \,dP(x)\biggr] & = \int \Ebb\Bigl[\Bigl(\wt{L}_{n,h}f_0(x)\Bigr)^2\Bigr] \,dP(x) \\
-& \leq \| T_{P,h}f_0 - f_0\|_P^2 + \frac{1}{n}\int \int \frac{1}{\bigl(d_{P,h}(x)\bigr)^2} \bigl(f_0(x') - f_0(x)\bigr)^2 \cdot \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr|^2 \,dP(x') \,dP(x) \\
+& \leq \| T_{P,h}f_0 - f_0\|_P^2 \\
+& \quad + \frac{1}{n}\int \int \frac{1}{\bigl(d_{P,h}(x)\bigr)^2} \bigl(f_0(x') - f_0(x)\bigr)^2 \cdot \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr|^2 \,dP(x') \,dP(x) \\
 & \leq \| T_{P,h}f_0 - f_0\|_P^2 + \frac{3h^2}{p_{\min}n} \wt{E}_{P,h}(f_0;\psi^2).
@@
 
-\paragraph{Proof of Lemma~\ref{lem:kernel_smoothing_bias}, $s = 1$.}
+\paragraph{Proof of Lemma~\ref{lem:kernel_smoothing_bias}, $s = 1$}
 Using the conclusions from Section~\ref{subsec:kernel_smoothing_preliminaries}, we have that
@@
 
-\paragraph{Proof of Lemma~\ref{lem:kernel_smoothing_bias}, $s > 1$.}
+\paragraph{Proof of Lemma~\ref{lem:kernel_smoothing_bias}, $s > 1$}
 Proceeding from~\eqref{pf:kernel_smoothing_bias_1}, we separate the integral into the portion sufficiently in the interior of $\mc{X}$ and that near the boundary, obtaining
@@
 
-\paragraph{Setup.}
+\paragraph{Setup}
 As usual, we assume the design points $(X_1,\ldots,X_n)$ are sampled independently from a distribution $P$, and the responses
@@
 
-\paragraph{Laplacian eigenmaps + kernel smoothing beats direct kernel smoothing.}
+\paragraph{Laplacian eigenmaps + kernel smoothing beats direct kernel smoothing}
 Now we are in a position to state our result. Let $P^{(n)}$ be the distribution with density $p^{(n)}$, i.e. the uniform probability distribution on $Q_1 \cup Q_2$. 
@@
 
-\paragraph{Proof of Proposition~\ref{prop:eigenmaps_beats_kernel_smoothing}.}
+\paragraph{Proof of Proposition~\ref{prop:eigenmaps_beats_kernel_smoothing}}
 First we prove~\eqref{eqn:eigenmaps_beats_kernel_smoothing_1}, then~\eqref{eqn:eigenmaps_beats_kernel_smoothing_2}.
@@
 
-\paragraph{A Useful Lemma.}
+\paragraph{A Useful Lemma}
 Let $m = 4n/\log^2n$. Furthermore, for $i = 0,\ldots,m - 1$, let
@@
 
-\paragraph{Proof (of Lemma~\ref{lem:balls_in_bins}).}
+\paragraph{Proof (of Lemma~\ref{lem:balls_in_bins})}
 For each $Q_{ij}$, we have that 

### 2021-10-16 manuscripts/laplacian_eigenmaps/appendix.tex
Updated Laplacian Eigenmaps manuscript.

@@
 \noindent
+
+\section{Upper bounds on population-level spectral series regression}
+
+In this section we first give the proof of Proposition~\ref{prop:spectral_series_estimation}, then of Proposition~\ref{prop:spectral_series_testing}.
+In both cases the structure of the analysis, which is fairly classical and straightforward, can be usefully compared to our analysis of PCR-LE (see Section~\ref{subsec:analysis}).
+
+\paragraph{Proof of Proposition~\ref{prop:spectral_series_estimation}.}
+We decompose risk into squared bias and variance,
+\begin{equation}
+\label{pf:spectral_series_estimation_0}
+\Ebb \|\wt{f} - f_0\|_P^2 = \Ebb\| \Ebb[\wt{f}]  - f_0\|_P^2 + \Ebb\| \wt{f} - \Ebb[\wt{f}]\|_P^2.
+\end{equation}
+Since the eigenfunctions $\{\psi_k\}$ form an orthonormal basis of $L^2(P)$, and $f_0 \in \mc{H}^{s}(\mc{X}) \subseteq L^2(P)$, we can write the squared bias in terms of squared Fourier coefficients of $f_0$, leading to the following upper bound,
+\begin{equation*}
+\|f_0 - \Ebb[\wt{f}]\|_P^2 = \sum_{k = K + 1}^{\infty}  \dotp{f_0}{\psi_k}_P^2 \leq  \frac{1}{\{\lambda_{K + 1}(\Delta_P)\}^s} \sum_{k = K + 1}^{\infty} \{\lambda_{k + 1}(\Delta_P)\}^s \dotp{f_0}{\psi_k}_P^2 \leq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}}{\{\lambda_{K + 1}(\Delta_P)\}^s}.
+\end{equation*}
+On the other hand, the variance term can be written as the sum of the variance of each empirical Fourier coefficient, and subsequently by the law of total variance we derive that
+\begin{align}
+\label{pf:spectral_series_estimation_2}
+\Ebb\| \wt{f} - \Ebb[\wt{f}]\|_P^2 = \sum_{k = 1}^{K} \Var\Bigl[\dotp{{\bf Y}}{\psi_k}_n\Bigr] & = \sum_{k = 1}^{K} \Var\Bigl[\Ebb[\dotp{Y}{\psi_k}_n|{\bf X}\Bigr] + \Ebb\Bigl[\Var[\dotp{Y}{\psi_k}_n|{\bf X}\Bigr] \nonumber \\
+& = \sum_{k = 1}^{K} \Var\Bigl[\dotp{f_0}{\psi_k}_n\Bigr] + \frac{1}{n}\Ebb\Bigl[\|\psi_k\|_n^2\Bigr] \nonumber \\
+& \leq \frac{K}{n} + \frac{1}{n}\sum_{k = 1}^{K}\Ebb\Bigl[\Bigl(f_0(X)\psi_k(X)\Bigr)^2\Bigr].
+\end{align}
+Consequently,
+\begin{equation}
+\label{pf:spectral_series_estimation_1}
+\Ebb \|\wt{f} - f_0\|_P^2 \leq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\bigl[\lambda_{K + 1}(\Delta_P)\bigr]^s} + \frac{K}{n} + \frac{1}{n}\Ebb\Bigl[(f_0(X))^2 \cdot \sum_{k = 1}^{K} (\psi_k(X))^2\Bigr].
+\end{equation}
+The claim of the proposition then follows from variants of two classical results in spectral geometry.
+The first is a Weyl's Law asymptotic scaling of the eigenvalues of $\Delta_P$ due to~\cite{dunlop2020}; formally, there exist constants $c$ and $C$ (which will depend on $P$ and $d$) such that
+\begin{equation}
+\label{eqn:weyl}
+ck^{2/d} \leq \rho_k \leq Ck^{2/d}\quad\textrm{for all $k \in \mathbb{N}$, $k \geq 2$}.
+\end{equation}
+The second is a local analog to Weyl's Law, which says that there exists a constant $C$ (again depending on $P$ and $d$) such that
+\begin{equation}
+\label{eqn:local_weyl}
+\sup_{x \in \mc{X}}\biggl\{\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2\biggr\} \leq CK \quad\textrm{for all $K \in \mathbb{N}$}.
+\end{equation}
+Equation~\eqref{eqn:local_weyl} is a direct implication of~\eqref{eqn:weyl} along with Theorem 17.5.3 of~\cite{hormander1973}.
+Plugging the upper bounds~\eqref{eqn:weyl} and~\eqref{eqn:local_weyl} back into~\eqref{pf:spectral_series_estimation_1}, we conclude that
+\begin{equation}
+\label{pf:spectral_series_estimation_3}
+\Ebb \|\wt{f} - f_0\|_P^2 \leq C\biggl(\frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{(K + 1)^{2s/d}} + \frac{K}{n}\biggr).
+\end{equation}
+If $n^{-1/2} \geq M$, then taking $K = 1$ implies $\Ebb \|\wt{f} - f_0\|_P^2 \leq C(M^2 + 1/n)$.
+Otherwise, setting $K = \floor{M^2n}^{d/(2s + d)}$ balances squared bias and variance, and yields the claim.
+\qed.
+
+\paragraph{Proof of Proposition~\ref{prop:spectral_series_testing}.}
+We briefly lay out the main ideas needed to prove Proposition~\ref{prop:spectral_series_testing}, following the lead of~\cite{ingster2009} who prove a similar result in the special case where $M = 1$ and $P$ is the uniform distribution over $\mc{X} = [0,1]^d$, and referring to that work for more details.
+
+We begin by computing the first two moments of the test statistic $\wt{T}$. The expectation is
+\begin{align*}
+\Ebb[\wt{T}] = \frac{(n - 1)}{n} \sum_{k = 1}^{K} \dotp{f_0}{\psi_K}_P^2 + \frac{K}{n} + \Ebb\Bigl[(f_0(X))^2 \dot \sum_{k = 1}^{K} (\psi_k(X))^2\Bigr],
+\end{align*}
+and from~\eqref{eqn:weyl} (Weyl's Law) we have that under the alternative $f_0 \neq 0$,
+\begin{equation*}
+\Ebb_{f_0}[\wt{T}] \geq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\lambda_{K + 1}^s(\Delta_{P})} + \frac{K}{n}.
+\end{equation*}
+To compute the variance, we decompose $\wt{T} = \wt{T}_{1,1} + \wt{T}_{1,2} + \wt{T}_{1,3} + \wt{T}_2$ into the sum of 3 U-statistics and the remaining diagonal terms, defined in terms of the equivalent kernel $\kappa(x,x') = \sum_{k = 1}^{K} \psi_k(x) \psi_k(x')$ as,
+\begin{align*}
+T_{1,1} & := \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} w_i w_j \kappa(X_i,X_j),\quad && T_{1,2} := \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} \bigl(w_i f_0(X_j) + w_jf_0(X_i)\bigr)\kappa(X_i,X_j) \\
+T_{1,3} & := \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} f_0(X_i) f_0(X_j) \kappa(X_i,X_j), \quad && T_2 := \frac{1}{n^2} \sum_{i = 1}^{n} Y_i^2 \kappa(X_i,X_i).
+\end{align*}
+The variances of each statistic can be found by routine computation (see~\cite{ingster2009}), and in particular satisfy the upper bounds
+\begin{align*}
+\Var(T_{1,1}) & \leq \frac{2K}{n^2}, \quad && \Var(T_{1,2}) \overset{\mathrm{(i)}}{\leq} \frac{C}{n}\|f_0\|_P^2\\
+\Var(T_{1,3}) & \overset{\mathrm{(ii)}}{\leq} C\biggl(\frac{K}{n}\|f_0\|_P^4 + \frac{K}{n^2}\|f_0\|_{L^4(\mc{X})}^4\biggr), \quad && \Var(T_{2}) \overset{\mathrm{(iii)}}{\leq} \frac{CK^2}{n^3}\biggl(1 + \|f_0\|_{L^4(\mc{X})}^4\biggr)
+\end{align*}
+where $\mathrm{(i)}-\mathrm{(iii)}$ hold due to local Weyl's law, i.e.~\eqref{eqn:local_weyl}. Upper bounds on Type I and Type II error,
+\begin{equation*}
+\Ebb_0[\wt{\varphi}] \leq \biggl(1 + CK/n^2\biggr)a, \quad \Ebb_{f_0}[1 - \wt{\varphi}] \leq \frac{C(K/n^2 + \|f_0\|_P^2 + K/n \|f_0\|_P^4 + K/n^2 \|f_0\|_{L^4(\mc{X})}^4)}{(\sum_{k = 1}^{K} \dotp{f_0}{\psi_k}_P^2 - \sqrt{2K/an})^2},
+\end{equation*}
+follow from Chebyshev's inequality. It can be verified that so long as
+\begin{equation}
+\label{pf:spectral_series_test}
+\|f_0\|_P^2 \geq C\Biggl(\frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\lambda_{K + 1}^s(\Delta_{P})} + \frac{\sqrt{K}}{n}\biggl(\sqrt{\frac{1}{a}} + \sqrt{\frac{1}{b}}\biggr)\Biggr)
+\end{equation}
+for a sufficiently large constant $C$, then $\Ebb_{f_0}[1 - \wt{\varphi}] \leq b$.
+The two summands in~\eqref{pf:spectral_series_test} are bias and standard deviation terms, respectively.
+When $M^2 \leq n^{-1}$, setting $K = 1$ gives the desired result.
+Otherwise, choosing $K = \floor{M^2n}^{2d/(4s + d)}$ balances these two terms, and leads to~\eqref{eqn:spectral_series_testing}.
+\qed
+% (AG 8/19/21): Same comment as above, regarding the role of M in random design regression testing.
 
@@
 
-\subsection{Upper bound on Estimation Error of Laplacian eigenmaps}
+\subsection{Upper bound on Estimation Error of PCR-LE}
 
@@
 	\label{lem:fixed_graph_estimation}
-	For any integer $s > 0$, and any integer $0 \leq K \leq n$, the Laplacian eigenmaps estimator $\wh{f}$ of~\eqref{eqn:laplacian_eigenmaps_estimator} satisfies
+	For any integer $s > 0$, and any integer $0 \leq K \leq n$, the PCR-LE estimator $\wh{f}$ of~\eqref{eqn:laplacian_eigenmaps_estimator} satisfies
 	\begin{equation}
@@
 
-\subsection{Upper bound on Testing Error of Laplacian Eigenmaps}
+\subsection{Upper bound on Testing Error of PCR-LE}
 
@@
 
-In Lemma~\ref{lem:approximation_error_nonlocal_laplacian_boundary}, we establish that if $f$ is Sobolev smooth of order $s > 2k$ and zero-trace, then near the boundary of $\mc{X}$ the non-local Laplacian $L_{P,\varepsilon}^kf$ is close to $0$ in the $\Leb^2$-sense.
+In Lemma~\ref{lem:approximation_error_nonlocal_laplacian_boundary}, we establish that if $f$ is Sobolev smooth of order $s > 2k$ and zero-trace, then near the boundary of $\mc{X}$ the non-local Laplacian $L_{P,\varepsilon}^kf$ is close to $0$ in the $L^2$-sense.
 \begin{lemma}
@@
 \begin{align}
-\|L_{P,\varepsilon}\|_{\Leb^2(U)}^2 & \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx \nonumber \\
+\|L_{P,\varepsilon}\|_{L^2(U)}^2 & \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx \nonumber \\
 & \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx.
@@
 \begin{equation*}
- E_{P,\varepsilon}(L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f;\mc{X}) \leq  \frac{C}{\varepsilon^2}\|L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f\|_{\Leb^2(\mc{X})}^2 \leq C |f|_{H^1(\mc{X})}^2.
+ E_{P,\varepsilon}(L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f;\mc{X}) \leq  \frac{C}{\varepsilon^2}\|L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f\|_{L^2(\mc{X})}^2 \leq C |f|_{H^1(\mc{X})}^2.
 \end{equation*}
@@
 	\label{eqn:gagliardo_nirenberg}
-	\|f\|_{\Leb^4(\mc{X})} \leq C_1 |f|_{H^s(\mc{X})}^{d/4s} \|f\|_{\Leb^2(\mc{X})}^{1 - d/(4s)} + C_2 \|f\|_{\Leb^2(\mc{X})}
+	\|f\|_{L^4(\mc{X})} \leq C_1 |f|_{H^s(\mc{X})}^{d/4s} \|f\|_{L^2(\mc{X})}^{1 - d/(4s)} + C_2 \|f\|_{L^2(\mc{X})}
 	\end{equation}
@@
 \begin{equation*}
-\frac{\Ebb[f^4(X)]}{\|f\|_P^4} \leq C \biggl(\frac{\|f\|_{\Leb^4(\mc{X})}}{\|f\|_{\Leb^2(\mc{X})}}\biggr)^4 \leq C_1\biggl(\frac{|f|_{H^s(\mc{X})}}{\|f\|_{\Leb^2(\mc{X})}}\biggr)^{d/s} + C_2,
+\frac{\Ebb[f^4(X)]}{\|f\|_P^4} \leq C \biggl(\frac{\|f\|_{L^4(\mc{X})}}{\|f\|_{L^2(\mc{X})}}\biggr)^4 \leq C_1\biggl(\frac{|f|_{H^s(\mc{X})}}{\|f\|_{L^2(\mc{X})}}\biggr)^{d/s} + C_2,
 \end{equation*}
@@
 \begin{equation*}
-C_1\biggl(\frac{|f|_{H^s(\mc{X})}}{\|f\|_{\Leb^2(\mc{X})}}\biggr)^{d/s} \leq \frac{\delta n}{64},
+C_1\biggl(\frac{|f|_{H^s(\mc{X})}}{\|f\|_{L^2(\mc{X})}}\biggr)^{d/s} \leq \frac{\delta n}{64},
 \end{equation*} 
@@
 	\label{eqn:gagliardo_nirenberg_manifold}
-	\|f\|_{\Leb^4(\mc{X})} \leq C_1 |f|_{H^s(\mc{X})}^{m/4s} \|f\|_{\Leb^2(\mc{X})}^{1 - m/(4s)} + C_2 \|f\|_{\Leb^2(\mc{X})}.
+	\|f\|_{L^4(\mc{X})} \leq C_1 |f|_{H^s(\mc{X})}^{m/4s} \|f\|_{L^2(\mc{X})}^{1 - m/(4s)} + C_2 \|f\|_{L^2(\mc{X})}.
 	\end{equation}
@@
 
-\section{Analysis of kernel smoothing}
-\label{subsec:kernel_smoothing}
-In this section we prove Lemma~\ref{lem:kernel_smoothing_insample} (in Section~\ref{subsec:pf_kernel_smoothing_insample}) and Lemma~\ref{lem:kernel_smoothing_bias} (in Section~\ref{subsec:pf_kernel_smoothing_bias}).
-In Section~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we give a sequence of design densities and regression functions $\{p^{(n)}(x), f_0^{(n)}(x)\}_{n \in \mathbb{N}}$ for which the estimator $T_{n,h}\wc{f}$ (extension of Laplacian eigenmaps by kernel smoothing) strictly outperforms directly kernel smoothing the responses, in the sense that
-\begin{equation*}
-\lim_{n \to \infty} \sup_{h'} \frac{\Ebb \|T_{n,h}\wh{f} - f_0\|_P^2}{\Ebb \|T_{h',n}Y - f_0\|_P^2 } = 0.
-\end{equation*}
-We begin with some preliminary estimates in Section~\ref{subsec:kernel_smoothing_preliminaries}, which will ease the subsequent analysis.
-\subsection{Some preliminary estimates}
-\label{subsec:kernel_smoothing_preliminaries}
-In certain parts the analysis of this section will overlap with Section~\ref{sec:graph_quadratic_form_euclidean}, where we upper bounded the non-local graph-Sobolev seminorm of a function $f$ in terms of the Sobolev norm of $f$.
-To see why this should be, note that for an function $f$ and point $x \in \mc{X}$, we have
-\begin{equation*}
-T_{P,h}f(x) - f(x) = \frac{1}{d_{Q,h}(x)} \int \bigl(f(x') - f(x)\bigr) \psi\biggl(\frac{\|x' - x\|}{h}\biggr) \,dQ(x') = \frac{h^{d + 2}}{d_{P,h}(x)} L_{P,h}f(x).
-\end{equation*}
-This expression reflects the known fact that the bias operator of kernel smoothing is equal to the non-local Laplacian, up to a rescaling by the population degree functional $d_{P,h}(x)$.
-In the second equality, we are using the notation $L_{P,h}f(x)$ exactly as defined in~\eqref{eqn:nonlocal_laplacian}, but with the kernel $\psi$ instead of $\eta$.
-Note that $\psi$ satisfies all the same assumptions as $\eta$, except that of positivity; when $\psi$ is a higher-order kernel it may take negative values.
-
-Now we provide a lower bound on $d_{P,h}(x)$ that holds uniformly over all $x \in \mc{X}$.
-Recall that by assumption the density $p$ is Lipschitz.
-Letting $L_p$ denote the Lipschitz constant of $p$, we have that
-\begin{align*}
-d_{P,h}(x) & = \int \psi\biggl(\frac{\|x' - x\|}{h}\biggr) p(x') \,dx' \\
-& = h^d \int \psi(\|z\|) p(hz + x) \1\Bigl\{ hz + x \in \mc{X} \Bigr\} \,dz \\
-& \geq h^d p(x) \int \psi(\|z\|) \1\Bigl\{ hz + x  \in \mc{X}\Bigr\} \,dz - L_p h^{d + 1} \|\psi\|_{\infty} \nu_d.
-\end{align*}
-Since by assumption $\mc{X}$ has Lipschitz boundary, setting $c_0$ to be a sufficiently small constant in~\ref{asmp:bandwidth}, we can further deduce that $\int \psi(\|z\|) \1\{hz + x \in \mc{X}\} \,dz \geq 1/3$, and consequently that
-\begin{equation}
-\label{eqn:degree_lower_bound}
-d_{P,h}(x) \geq \frac{p(x)}{3} h^d \geq \frac{p_{\min}}{3}h^d \quad \textrm{for all $x \in \mc{X}$.}
-\end{equation}
-
-\subsection{Proof of Lemma~\ref{lem:kernel_smoothing_insample}}
-\label{subsec:pf_kernel_smoothing_insample}
-
-To begin with, we apply the triangle inequality to upper bound $\|T_{n,h}\wc{f} - f_0\|_P$ by the sum of two terms,
-\begin{equation}
-\label{pf:kernel_smoothing_insample_1}
-\|T_{n,h}\wc{f} - f_0\|_P \leq \|T_{n,h}(\wc{f} - f_0)\|_P + \|T_{n,h}f_0 - f_0\|_P.
-\end{equation}
-We proceed by separately upper bounding each term on the right hand side of~\eqref{pf:kernel_smoothing_insample_1}. We will show that
-\begin{equation}
-\label{pf:kernel_smoothing_insample_2}
-\|T_{n,h}(\wc{f} - f_0)\|_P^2 \leq C \|\wc{f} - f_0\|_n^2
-\end{equation}
-and that 
-\begin{equation}
-\label{pf:kernel_smoothing_insample_3}
-\|T_{n,h}f_0 - f_0\|_P^2 \leq \frac{C}{\delta} \cdot \frac{h^2}{nh^d} |f|_{H^1(\mc{X})}^2 + \frac{C}{\delta}\|T_{h,P}f_0 - f_0\|_P^2,
-\end{equation}
-each with probability at least $1 - C\exp(-cnh^d)$. Together these will imply the claim.
-
-\underline{\emph{Proof of~\eqref{pf:kernel_smoothing_insample_2}}.}
-Fix $x \in \mc{X}$. By the Cauchy-Schwarz inequality we have
-\begin{align*}
-\Bigl[T_{n,h}\bigl(\wc{f} - f_0\bigr)(x)\Bigr]^2 & = \Biggl[\frac{1}{d_{n,h}(x)^2}\int \psi\biggl(\frac{\|x' - x\|}{h}\biggr) \cdot \bigl(\wc{f}(x') - f_0(x')\bigr) \,dP_n(x')\Biggr]^2 \\
-& \leq \Biggl[\frac{1}{d_{n,h(x)}^2} \int \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr| \,dP_n(x')\Biggr] \cdot \Biggl[\int \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr| \cdot \bigl(\wc{f}(x') - f_0(x')\bigr)^2 \,dP_n(x')\Biggr] \\
-& = \frac{d_{n,h}^{+}(x)}{|d_{n,h}(x)|} \cdot \frac{1}{|d_{n,h}(x)|} \Biggl[\int \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr| \cdot \bigl(\wc{f}(x') - f_0(x')\bigr)^2 \,dP_n(x')\Biggr].
-\end{align*}
-In the last line all we have done is written $d_{n,h}^{+}(x)$ for the degree functional computed with respect to the kernel $|\psi|$, recalling that $\psi$ may take negative values so $d_{n,h}^{+}(x)$ may not be equal to $d_{n,h}(x)$.
-
-Now we integrate over $x \in \mc{X}$ to get 
-\begin{align}
-\|T_{n,h}(\wc{f} - f_0)\|_P^2 & = \int \Bigl[T_{n,h}\bigl(\wc{f} - f_0\bigr)(x)\Bigr]^2 \,dP(x) \nonumber \\
-& \leq \int \int \frac{d_{n,h}^{+}(x)}{|d_{n,h}(x)|} \cdot \frac{1}{|d_{n,h}(x)|} \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr| \cdot \bigl(\wc{f}(x') - f_0(x')\bigr)^2 \,dP_n(x') \,dP(x) \nonumber \\
-& \leq \sup_{x \in \mc{X}} \frac{d_{n,h}^{+}(x)}{|d_{n,h}(x)|} \cdot \int \int  \frac{1}{|d_{n,h}(x)|} \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr| \cdot \bigl(\wc{f}(x') - f_0(x')\bigr)^2 \,dP(x) \,dP_n(x') \nonumber \\
-& \leq \sup_{x \in \mc{X}} \frac{d_{n,h}^{+}(x) d_{P,h}^+(x)}{|d_{n,h}(x)|^2} \cdot \|\wc{f} - f_0\|_n^2. \label{pf:kernel_smoothing_insample_4}
-\end{align}
-Thus we have reduced the problem to showing that the various degree functionals $d_{n,h}^{+}, d_{P,h}^{+}$ and $d_{n,h}$ all put similar weight on a given point $x$.
-We use~\eqref{eqn:uniform_bound_empirical_degree_2}, which gives a uniform multiplicative bound on deviations of the empirical degree around its mean, to conclude that with probability at least $1 - C\exp\{-cnh^d\}$,
-\begin{equation*}
-d_{n,h}(x) \geq \frac{1}{2}d_{P,h}(x)~~ \textrm{and} ~~ d_{n,h}^{+}(x) \leq \frac{3}{2} d_{P,h}^{+}(x) \quad\textrm{for all $x \in \mc{X}$.}
-\end{equation*}
-Therefore,
-\begin{equation*}
-\sup_{x \in \mc{X}} \frac{d_{n,h}^{+}(x) d_{P,h}^+(x)}{|d_{n,h}(x)|^2} \leq 6 \cdot \sup_{x \in \mc{X}} \frac{|d_{P,h}^{+}(x)|^2}{|d_{P,h}(x)|^2} \leq 36 \biggl(\frac{\|\psi\|_{\infty} p_{\max} \nu_d}{p_{\min}}\biggr)^2
-\end{equation*}
-with the second inequality following from~\eqref{eqn:degree_lower_bound}. Plugging this back into~\eqref{pf:kernel_smoothing_insample_4} gives the claim.
-
-\underline{\emph{Proof of~\eqref{pf:kernel_smoothing_insample_3}}.}
-At a given point $x \in \mc{X}$, we have
-\begin{align*}
-T_{n,h}f_0(x) - f_0(x) & = \frac{1}{d_{n,h}(x)} \sum_{i = 1}^{n} \bigl(f_0(X_i) - f_0(x)\bigr) \psi\biggl(\frac{\|X_i - x\|}{h}\biggr) \\
-& = \frac{d_{P,h}(x)}{d_{n,h}(x)} \cdot \frac{1}{nd_{P,h}(x)} \sum_{i = 1}^{n} \bigl(f_0(X_i) - f_0(x)\bigr) \psi\biggl(\frac{\|X_i - x\|}{h}\biggr).
-\end{align*}
-Thus,
-\begin{equation}
-\label{pf:kernel_smoothing_insample_5}
-\Bigl[T_{n,h}f_0(x) - f_0(x)\Bigr]^2 = \biggl[\frac{d_{P,h}(x)}{d_{n,h}(x)}\biggr]^2 \cdot \biggl[\underbrace{\frac{1}{nd_{P,h}(x)}\sum_{i = 1}^{n} \bigl(f_0(X_i) - f_0(x)\bigr) \psi\biggl(\frac{\|X_i - x\|}{h}\biggr)}_{:= \wt{L}_{n,h}f_0(x)}\biggr]^2
-\end{equation}
-In the proof of~\eqref{pf:kernel_smoothing_insample_2} we have already given an upper bound on the ratio of population to empirical degree, which implies that
-\begin{equation*}
-\sup_{x \in \mc{X}} \biggl[\frac{d_{P,h}(x)}{d_{n,h}(x)}\biggr]^2 \leq 4,
-\end{equation*}
-with probability at least $1 - C\exp\{-cnh^d\}$. On the other hand, we note that the second term in the product in~\eqref{pf:kernel_smoothing_insample_5} has expectation
-\begin{equation*}
-\Ebb\Bigl[\wt{L}_{n,h}f_0(x)\Bigr] = T_{P,h}f_0(x) - f_0(x),
-\end{equation*} 
-and variance 
-\begin{equation*}
-\Var\Bigl[\wt{L}_{n,h}f_0(x)\Bigr] \leq \frac{1}{n(d_{P,h}(x))^2} \Ebb\biggl[(f_0(X) - f_0(x))^2 \cdot \biggl|\psi\biggl(\frac{\|X - x\|}{h}\biggr)\biggr|^2\biggr].
-\end{equation*}
-Integrating with respect to $P$ gives
-\begin{align*}
-\Ebb\biggl[\int \Bigl(\wt{L}_{n,h}f_0(x)\Bigr)^2 \,dP(x)\biggr] & = \int \Ebb\Bigl[\Bigl(\wt{L}_{n,h}f_0(x)\Bigr)^2\Bigr] \,dP(x) \\
-& \leq \| T_{P,h}f_0 - f_0\|_P^2 + \frac{1}{n}\int \int \frac{1}{\bigl(d_{P,h}(x)\bigr)^2} \bigl(f_0(x') - f_0(x)\bigr)^2 \cdot \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr|^2 \,dP(x') \,dP(x) \\
-& \leq \| T_{P,h}f_0 - f_0\|_P^2 + \frac{3h^2}{p_{\min}n} \wt{E}_{P,h}(f_0;\psi^2).
-\end{align*}
-In the final inequality we have used the lower bound on $d_{P,h}(x)$ from~\eqref{eqn:degree_lower_bound}, and written $E_{P,h}(f_0;\psi^2)$ for the non-local Dirichlet energy defined with respect to the kernel $\psi^2$.
-
-Putting the pieces together, we conclude that
-\begin{align*}
-\|T_{n,h}f_0(x) - f_0(x)\|_P^2 & = \int \bigl(T_{n,h}f_0(x) - f_0(x)\bigr)^2 \,dP(x) \\
-& \leq \sup_{x \in \mc{X}} \biggl[\frac{d_{P,h}(x)}{d_{n,h}(x)}\biggr]^2 \cdot \int \Bigl(\wt{L}_{n,h}f_0(x)\Bigr)^2 \,dP(x) \\
-& \overset{(i)}{\leq} 4\frac{\| T_{P,h}f_0 - f_0\|_P^2}{\delta} + \frac{12h^2}{\delta p_{\min}nh^d} {E}_{P,h}(f_0;\psi^2) \\
-& \overset{(ii)}{\leq}  4\frac{\| T_{P,h}f_0 - f_0\|_P^2}{\delta} + \frac{Ch^2}{\delta p_{\min}nh^d} |f_0|_{H^1(\mc{X})}^2,
-\end{align*}
-with probability at least $1 - \delta - C\exp(-cnh^d)$.
-In $(i)$ we have used Markov's inequality, and in $(ii)$ we have applied the estimate~\eqref{pf:estimate_nonlocal_seminorm_1} to the non-local Dirichlet energy ${E}_{P,h}(f_0;\psi^2)$.
-This establishes~\eqref{pf:kernel_smoothing_insample_3}.
-
-\subsection{Kernel smoothing bias}
-\label{subsec:pf_kernel_smoothing_bias}
-Lemma~\ref{lem:kernel_smoothing_bias} gives the necessary upper bounds on the bias of kernel smoothing.
-\begin{lemma}
-	\label{lem:kernel_smoothing_bias}
-	Suppose Model~\ref{def:model_flat_euclidean}, and that the kernel smoothing operator $T_{P,h}$ is computed with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel}.
-	\begin{itemize}
-		\item If $f_0 \in H^1(\mc{X})$, then there exists a constant $C$ which does not depend on $f_0$ such that
-		\begin{equation*}
-		\|T_{P,h}f_0 - f_0\|_P^2 \leq C h^{2} |f|_{H^1(\mc{X})}^2.
-		\end{equation*}
-		\item If $f_0 \in H_0^{s}(\mc{X})$, $p \in C^{s - 1}(\mc{X})$, and $\eta$ satisfies~\ref{asmp:ho_kernel}, then there exists a constant $C$ which does not depend on $f_0$ such that
-		\begin{equation*}
-		\|T_{P,h}f_0 - f_0\|_P^2 \leq C h^{2s} |f|_{H^s(\mc{X})}^2.
-		\end{equation*}
-	\end{itemize}
-\end{lemma}
-We separately prove the first-order ($s = 1$) and higher-order ($s > 1$) parts of Lemma~\ref{lem:kernel_smoothing_bias}.
-In both cases, the proof will rely heavily on results already established regarding the non-local Laplacian $L_{P,h}$ and non-local Dirichlet energy $E_{P,h}$, which we recall are given for a kernel function $\mc{K}$ by
-\begin{equation*}
-L_{P,h}f(x) = \frac{1}{h^{d + 2}} \int \bigl(f(x') - f(x)\bigr)\mc{K}\biggl(\frac{\|x' - x\|}{h}\biggr)\,dP(x'),
-\end{equation*}
-and $E_{P,h}(f;\mc{K}) = \dotp{L_{P,h}f}{f}_{P}$, respectively.
-
-\paragraph{Proof of Lemma~\ref{lem:kernel_smoothing_bias}, $s = 1$.}
-Using the conclusions from Section~\ref{subsec:kernel_smoothing_preliminaries}, we have that
-\begin{equation}
-\label{pf:kernel_smoothing_bias_1}
-\|T_{P,h}f - f\|_P^2 \leq \frac{9h^4}{p_{\min}^2} \int \bigl[L_{P,h}f(x)\bigr]^2 \,dP(x).
-\end{equation}
-By the Cauchy-Schwarz inequality, we have that
-\begin{align*}
-\int \bigl[L_{P,h}f(x)\bigr]^2 \,dP(x) & = \frac{1}{h^{2d + 4}}\int \biggl[\int \bigl(f(x') - f(x)\bigr)\psi\biggl(\frac{\|x' - x\|}{h}\biggr) \,dP(x')\biggr]^2 \,dP(x) \\
-& \leq \frac{C}{h^{d + 4}} \int \int \bigl(f(x') - f(x)\bigr)^2 \cdot \biggl|\psi\biggl(\frac{\|x' - x\|}{h}\biggr)\biggr| \,dP(x') \,dP(x) \\
-& = \frac{C}{h^{2}} E_{P,h}(f;|\psi|).
-\end{align*}
-Applying the estimate~\eqref{pf:estimate_nonlocal_seminorm_1} to the non-local Dirichlet energy ${E}_{P,h}(f;|\psi|)$ and plugging back into~\eqref{pf:kernel_smoothing_bias_1} gives the claimed result.
-
-\paragraph{Proof of Lemma~\ref{lem:kernel_smoothing_bias}, $s > 1$.}
-Proceeding from~\eqref{pf:kernel_smoothing_bias_1}, we separate the integral into the portion sufficiently in the interior of $\mc{X}$ and that near the boundary, obtaining
-\begin{equation}
-\label{pf:kernel_smoothing_bias_2}
-\|T_{P,h}f - f\|_P^2 \leq\frac{9p_{\max}h^4}{p_{\min}^2}\Bigl(\|L_{P,h}f\|_{L^2(\mc{X}_h)}^2 + \|L_{P,h}f\|_{L^2(\partial_{h}(\mc{X}))}^2\Bigr).
-\end{equation}
-In Lemma~\ref{lem:approximation_error_nonlocal_laplacian_boundary}, we established a sufficient upper bound on the second term,
-\begin{equation*}
-\|L_{P,h}f\|_{L^2(\partial_{h}(\mc{X}))}^2 \leq Ch^{2(s - 2)} \|f\|_{H^s(\mc{X})}^2.
-\end{equation*}
-Thus it remains to upper bound the first term. Here we recall that at a given $x \in \mc{X}_h$, we can write
-\begin{align*}
-L_{P,h}f(x) & = \frac{1}{h^{2}}\sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1}\frac{h^{j_1 + j_2}}{j_1!j_2!}  \int d_x^{j_1}f(z) d_x^{j_2}p(z) \psi\bigl(\|z\|\bigr) \,dz \quad + \\
-& \quad \frac{1}{h^{2}} \sum_{j = 1}^{s - 1} \frac{h^j}{j!} \int d_x^jf(z)  r_{zh + x}^{q}(x;p) \psi\bigl(\|z\|\bigr) \,dz \quad  + \\
-& \quad \frac{1}{h^{2}} \int r_{zh + x}^j(x;f) \psi\bigl(\|z\|\bigr) p(zh + x)\,dz \\
-& = G_1(x) + G_2(x) + G_3(x).
-\end{align*}
-(Here $q = s - 1$.)
-
-We have already given sufficient upper bounds on $\|G_j\|_{L^2(\mc{X}_{h})}$ for $j = 2,3$ in~\eqref{pf:approximation_error_nonlocal_laplacian_2}.
-Thus it remains only to upper bound $\|G_1\|_{L^2(\mc{X}_h)}$.
-Recall the expansion of $G_1$ from~\eqref{pf:approximation_error_nonlocal_laplacian_3},
-\begin{equation*}
-G_1(x) = \sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1} \frac{h^{j_1 + j_2 - 2}}{j_1!j_2!}  \underbrace{\int_{B(0,1)} d_x^{j_1}f(z) d_x^{j_2}p(z) \eta(\|z\|) \,dz}_{:= g_{j_1,j_2}(x)}.
-\end{equation*}
-Noting that $d_x^{j_1} \cdot d_x^{j_2}$ is a degree-$(j_1 + j_2)$ multivariate polynomial, and recalling that $\psi$ is an order-$s$ kernel, we have that 
-\begin{equation*}
-\int g_{j_1,j_2}(z) \psi\bigl(\|z\|\bigr) \,dz = 0,\quad \textrm{for all $j_1,j_2$ such that $j_1 + j_2 < s$.}
-\end{equation*}
-Otherwise, derivations similar to those used in the proof of Lemma~\ref{lem:approximation_error_nonlocal_laplacian} imply that
-\begin{equation*}
-\|g_{j_1,j_2}\|_{L^2(\mc{X}_h)} \leq C \|f\|_{H^s(\mc{X})} \|p\|_{C^{s - 1}(\mc{X})},\quad \textrm{for all $j_1,j_2$ such that $j_1 + j_2 \geq s$,}
-\end{equation*}
-from which it follows that
-\begin{equation*}
-\|G_1\|_{L^2(\mc{X}_h)}^2 \leq C  h^{2(s - 2)} \|f\|_{H^s(\mc{X})} \|p\|_{C^{s - 1}(\mc{X})}.
-\end{equation*}
-Together these upper bounds on $\|G_j\|_{L^2(\mc{X}_j)}$ for $j = 1,2,3$ imply that
-\begin{equation*}
-\|L_{P,h}f\|_{\mc{X}_h}^2 \leq Ch^{2(s - 2)} \|f\|_{H^s(\mc{X})}^2,
-\end{equation*}
-and plugging this back into~\eqref{pf:kernel_smoothing_bias_2} yields the claim.
-
-\subsection{Comparison of Laplacian Eigenmaps to kernel smoothing}
+\section{Graph Laplacian methods and the cluster assumption}
 \label{subsec:eigenmaps_beats_kernel_smoothing}
 
-In this section, we give a sequence of densities $p^{(n)}$ and regression functions $f_0^{(n)}$ for which the $L^2(P)$ error of the two-stage estimator $T_{n,h}\wh{f}$ is stochastically dominated by the $L^2(P)$ error of applying kernel smoothing directly to the responses.
-We begin by precisely setting up the problem and reviewing the two estimators we will compare.
-Then we give and prove our formal claim.
-
-\paragraph{Setup.}
-As usual, we assume the design points $(X_1,\ldots,X_n)$ are sampled independently from a distribution $P$, and the responses
-\begin{equation*}
-Y_i = f_0(X_i) + \varepsilon_i,
-\end{equation*}
-where the noise terms $\varepsilon_i \sim N(0,1)$ are sampled independently, and are also independent of the design points.
-
-Unlike in our main text, we shall assume a specific (parametric) form for $p = p^{(n)}$ and $f_0 = f_0^{(n)}$, that changes as a function of $n$. More specifically, using the notation
-\begin{equation*}
-r := \frac{(\log n)^2}{n}, \quad Q_1 := [0,1/2 - r], \quad Q_2 := [1/2 + r,1],
-\end{equation*}
-we take the domain $\mc{X}^{(n)} = Q_1 \cup Q_2$, and 
+A main conclusion of our paper is that PCR-LE is minimax optimal for nonparametric regression over certain Sobolev classes.
+It is not the only optimal method.
+For instance, kernel smoothing and least squares using an appropriate set of basis functions as features are two other minimax optimal methods over these Sobolev classes.
+We now give an example where PCR-LE is better than these two alternatives, in the sense of having (much) smaller risk.
+This is possible because PCR-LE performs remarkably well when the regression function $f_0$ and design distribution $P$ satisfy a \emph{cluster assumption}: that is, when the regression function is (approximately) piecewise constant over high-density clusters of the design distribution $P$.
+On the other hand, kernel smoothing (with Euclidean distance) and least squares (using eigenfunctions of an unweighted Laplace operator) cannot take advantage of the cluster assumption.
+We call this property of PCR-LE \emph{density adaptivity}.
+
+\subsection{Setup}
+We begin by specifying a sequence of design densities and regression functions $\{(p^{(n)}, f_0^{(n)}): n \in \mathbb{N}\}$.
+These distributions will all be chosen to satisfy the cluster assumption.
+To that end, we define two clusters $Q_1,Q_2 \subset \Reals$ using a cluster separation parameter $r$, as
+\begin{equation*}
+Q_1 := [0,1/2 - r], \quad Q_2 := [1/2 + r,1],
+\end{equation*}
+and take the domain $\mc{X}^{(n)} := Q_1 \cup Q_2$.
+We then take the design density to be uniform over $\mc{X}^{(n)}$ and the regression function to be a piecewise constant function over $Q_1$ and $Q_2$ of height $\theta$,
 \begin{equation}
 \label{def:model_cluster_assumption}
-p^{(n)}(x) := \frac{1}{1 - 2r}\1\bigl\{x \in Q_1 \cup Q_2\bigr\}, \quad f_0^{(n)}(x) := \theta \cdot \Bigl(\1\bigl\{x \in Q_1\bigr\} - \1\bigl\{x \in Q_2\bigr\}\Bigr),
-\end{equation}
-where $\theta \in \Reals$.
-Note that as $n \to \infty$, the density $p^{(n)}$ is converging (in, say, $L^2([0,1])$ norm) to the uniform distribution on $[0,1]$, and the regression function $f_0^{(n)}$ to a step function.
-
-Our goal is to estimate $f_0$ using an estimator $f$, in such a way that the resulting $L^2(P)$ error $\|f - f_0\|_P^2$ is small.
-We shall consider two estimators: Laplacian eigenmaps + extension by kernel smoothing, and direct kernel smoothing of the responses.
-In the notation of our main text, these are respectively $T_{n,h}\wh{f}$ and $T_{n,h'}{\bf Y}$.
-For simplicity, in both cases we shall consider only the boxcar kernel,
+p^{(n)}(x) := \frac{1}{1 - 2r}\1\bigl\{x \in Q_1 \cup Q_2\bigr\}, \quad f_0^{(n)}(x) := \theta \cdot \Bigl(\1\bigl\{x \in Q_1\bigr\} - \1\bigl\{x \in Q_2\bigr\}\Bigr).
+\end{equation}
+Thus $p^{(n)}$ and $f_0^{(n)}$ belong to a two-parameter family, where the parameters are the cluster separation $r$ and height $\theta$.
+Generally speaking, the smaller the separation $r$, and the larger the height $\theta$, the more graph Laplacian methods will outperform both kernel smoothing and linear regression using eigenfunctions of the unweighted Laplace operator as features.
+
+We now define kernel smoothing and least squares using eigenfunction of an unweighted Laplace operator For a kernel function $\psi$ and bandwidth parameter $h$, the kernel smoothing estimator $\wt{f}_{\mathrm{KS}}$ is defined at a point $x \in \mc{X}$ as
+\begin{equation}
+\label{eqn:kernel_smoothing}
+\wt{f}_{\mathrm{KS}}(x) := 
+\begin{dcases*}
+0, & \quad \textrm{if $d_{n,h}(x) = 0$,} \\
+\frac{1}{d_{n,h}(x)} \sum_{i = 1}^{n} Y_i \psi\biggl(\frac{\|X_i - x\|}{h}\biggr), & \quad \textrm{otherwise.}
+\end{dcases*}
+\end{equation}
+
+Let $(\lambda_1,\phi_1), (\lambda_2,\phi_2),\ldots$ be eigenpairs of the unweighted Laplace operator $\Delta$ on $[0,1]$, meaning
+\begin{equation}
+\label{eqn:laplace_beltrami_eigenpairs}
+\Delta \phi_k = \lambda_k \phi_k, ~~ \|\phi_k\|_{L^2([0,1])} = 1,~~\frac{d}{dx}\phi_k(0) = \frac{d}{dx}\phi_k(1) = 0. 
+\end{equation}
+In this case the eigenfunctions $\phi_k$ of $\Delta$ are simply cosine functions, with eigenvalues proportional to their squared frequency.
+Noting that $\phi_1(x) = 1$ and $\lambda_1 = 0$, for $k = 2,3,\ldots$ we have
+\begin{equation*}
+\phi_k(x) = \sqrt{2}\cdot\cos(2\pi k x),~~\lambda_k(\Delta) = \pi^2 k^2.
+\end{equation*}
+The least squares estimator using $\phi_1,\ldots,\phi_K$ ($1 \leq K \leq n$) eigenfunctions as features is simply\footnote{For convenience, we will assume $\Phi \in \Reals^{n \times K}$ is full rank.
+If this is not the case, the least squares estimator $\wt{f}_K$ is not uniquely defined, but any solution will equal ${\bf Y}$ in-sample, and will satisfy $\|\wt{f}_K - f_0\|_n^2 \geq 1/2$ with high probability.}
+\begin{equation}
+\label{eqn:least_squares}
+\wt{f}_{K} := \argmin_{f \in \mathrm{span}\{\phi_1,\ldots,\phi_K\}} \|Y - f\|_n^2 = \Phi (\Phi^{\top} \Phi)^{-1} \Phi^{\top} Y.
+\end{equation}
+Hereafter, we will refer to $\wt{f}_K$ as the \emph{uniform least squares} estimator.
+
+\subsection{Upper bounds on risk of PCR-LE}
+Now we are in a position to state our results. Both PCR-LE and kernel smoothing depend in part on the choice of kernel. For simplicity, in our analysis we only consider the boxcar kernel,
 \begin{equation}
@@
 \end{equation}
-
-\paragraph{Laplacian eigenmaps + kernel smoothing beats direct kernel smoothing.}
-Now we are in a position to state our result. Let $P^{(n)}$ be the distribution with density $p^{(n)}$, i.e. the uniform probability distribution on $Q_1 \cup Q_2$. 
+This is strictly for convenience, and the following results will also hold for any kernel that satisfies~\ref{asmp:kernel_flat_euclidean}.
+
 \begin{proposition}
-	\label{prop:eigenmaps_beats_kernel_smoothing}
+	\label{prop:cluster_assumption_ub}
 	Suppose $(X_1,Y_1),\ldots(X_n,Y_n)$ are sampled according to~\eqref{def:model_cluster_assumption}.
+Compute the PCR-LE estimator $\wh{f}$ using a kernel $\eta$ which satisfies~\eqref{asmp:boxcar_kernel}, number of eigenvectors $K = 2$, and radius $\varepsilon = r/2$.
+Then,
+	\begin{equation}
+	\label{eqn:cluster_assumption_le_ub}
+	\Ebb\Bigl[\|\wh{f} - f_0^{(n)}\|_{n}^2\Bigr] \leq \biggl(6\theta^2 + \frac{1}{n}\biggr) \cdot \frac{8}{r} \exp(-nr/8) + \frac{1}{n}.
+	\end{equation}
+\end{proposition}
+\paragraph{Proof of Proposition~\ref{prop:cluster_assumption_ub}.}
+
+We begin by showing that, with high probability, the eigenvectors $v_1,v_2$ respect the cluster structure of $p^{(n)}$.
+Denote $u_1 = (\1\{X_i \in Q_1\})_{i \in [n]}$, and likewise $u_2 = (\1\{X_i \in Q_2\})_{i \in [n]}$.
+We make the following two observations:
+\begin{enumerate}
+	\item Because $\varepsilon < r$ and the kernel $\eta$ is compactly supported on $[0,1]$, for each $X_i \in Q_1$ and $X_j \in Q_2$, it must be the case that $\eta(\|X_i - X_j\|/\varepsilon) = 0$.
+	\item Using an elementary concentration argument (stated in Lemma~\ref{lem:balls_in_bins}) and the triangle inequality, we deduce that with probability at least $1 - 4/\varepsilon\exp(-n\varepsilon/4)$ there exists a path in $G_{n,\varepsilon}$ between each $X_i, X_j \in Q_1$, and likewise between each $X_i,X_j \in Q_2$.
+\end{enumerate}
+Together these observations imply that with high probability the neighborhood graph $G_{n,\varepsilon}$ consists of exactly two connected components: one consisting of all design points $X_i \in Q_1$, and the other consisting of all design points $X_i \in Q_2$.
+In other words,
+\begin{equation}
+\label{pf:eigenmaps_beats_kernel_smoothing_1}
+\Pbb\Bigl(\mathrm{span}\{v_1,v_2\} = \mathrm{span}\{u_1,u_2\}\Bigr) \geq 1 - 4/\varepsilon\exp(-n\varepsilon/4).
+\end{equation}
+
+Let us condition on the ``good'' event $\mc{E}$ that the design points $X_1,\ldots,X_n$ satisfy~\eqref{eqn:balls_in_bins}, and therefore that $\mathrm{span}\{v_1,v_2\} = \mathrm{span}\{u_1,u_2\}$.
+Consider the empirical mean $\wb{Y}_Q := \frac{1}{\sharp\{Q \cup {\bf X}\}} \sum_{i: X_i \in Q} Y_i$.
+Since $\mathrm{span}\{v_1,v_2\} = \mathrm{span}\{u_1,u_2\}$, the estimator $\wh{f} = \wh{f}_{\mathrm{LE}}$ will be piecewise constant on $Q_1$ and $Q_2$, and in fact we have that
+\begin{equation}
+\label{pf:eigenmaps_beats_kernel_smoothing_-1}
+\wh{f} = \wb{Y}_{Q_1}u_1 + \wb{Y}_{Q_2}u_2.
+\end{equation}
+Therefore conditional on $\mc{E}$,
+\begin{equation*}
+\|\wh{f} - f_0^{(n)}\|_{n}^2 = P_n(Q_1) \cdot (\wb{Y}_{Q_1} - \theta)^2 + P_n(Q_2) \cdot (\wb{Y}_{Q_2} + \theta)^2
+\end{equation*}
+and consequently,
+\begin{equation}
+\label{pf:eigenmaps_beats_kernel_smoothing_0}
+\Ebb\Bigl[\|\wh{f} - f_0^{(n)}\|_{n}^2 \Big|\mc{E}\Bigr] = \Ebb\Bigl[\Ebb\bigl[\|\wh{f} - f_0^{(n)}\|_{n}^2\big|X_1,\ldots,X_n\bigr] \Bigr|\mc{E}\Bigr] = \frac{1}{n}.
+\end{equation}
+
+Now we derive a crude upper bound on $\|\wh{f} - f_0^{(n)}\|_{n}$ that will suffice to control the error conditional on $\mc{E}^c$. We observe that the empirical norm of $\wh{f}$ is bounded,
+\begin{equation*}
+\|\wh{f}\|_n^2 \leq \frac{2}{n}\sum_{i = 1}^{n}\dotp{Y}{v_1}_n^2v_{1,i}^2 + \dotp{{\bf Y}}{v_2}_n^2v_{2,i}^2 \leq 2\bigl(\dotp{{\bf Y}}{v_1}_n^2 + \dotp{{\bf Y}}{v_2}_n^2\bigr) \leq 4 \|{\bf Y}\|_n^2.
+\end{equation*}
+Noting that $\Ebb[\|{\bf Y}\|_n^2|X_1,\ldots,X_n] = \|f_0\|_n^2 + 1/n  = \theta^2 + 1/n$, we conclude that
+\begin{equation*}
+\Ebb\Bigl[\|\wh{f} - f_0\|_{n}^2 \cdot \1\{\mc{E}^c\}\Bigr] \leq \Ebb\Bigl[\Bigl(2 \|f_0\|_{n}^2 + 4(\theta^2 + 1/n) \cdot \1\{\mc{E}^c\} \Bigr] \leq (6\theta^2 + n^{-1}) \cdot 4\varepsilon^{-1} \exp(-n\varepsilon/4).
+\end{equation*}
+Combining this with~\eqref{pf:eigenmaps_beats_kernel_smoothing_0} implies~\eqref{eqn:cluster_assumption_le_ub}.
+
+\subsection{Lower bounds on risk of kernel smoothing and least squares}
+
+\begin{proposition}
+	\label{prop:cluster_assumption_lb}
+	Suppose $(X_1,Y_1),\ldots,(X_n,Y_n)$ are sampled according to~\eqref{def:model_cluster_assumption}. Suppose $(\log n)^2/n \leq r \leq c$, where $c$ is a universal constant. 
 	\begin{itemize}
-		\item Compute the Laplacian eigenmaps estimator $\wh{f}$ using a kernel $\eta$ which satisfies~\eqref{asmp:boxcar_kernel}, number of eigenvectors $K = 2$, and radius $\varepsilon = ((\log n)^2/n)/2$.
-Compute the kernel smoothed extension $T_{n,h}\wh{f}$ using a kernel $\psi$ which satisfies~\eqref{asmp:boxcar_kernel}, and bandwidth $h = ((\log n)^2/n)/2$.
-Then,
-		\begin{equation}
-		\label{eqn:eigenmaps_beats_kernel_smoothing_1}
-		\Ebb \|T_{n,h}\wh{f} - f_0^{(n)}\|_{L^2(P^{(n)})}^2 \leq \frac{8 (\log n)^2 + 6 (\log n)^2 \theta^2}{n}.
-		\end{equation}
-		\item Compute the kernel smoothing estimator $T_{n,h}{\bf Y}$, using a kernel $\psi$ which satisfies~\eqref{asmp:boxcar_kernel}. Then
-		\begin{equation}
-		\label{eqn:eigenmaps_beats_kernel_smoothing_2}
-		\inf_{h' > 0} \Ebb\Bigl[\|T_{n,h'}{\bf Y} - f_0^{(n)}\|_{L^2(P^{(n)})}^2\Bigr] \geq \min\biggl\{ \frac{1}{16(\log n)^2}, \frac{\theta}{\sqrt{32 n}} \biggr\}.
+		\item Compute the kernel smoothing estimator $\wt{f} = \wt{f}_{\mathrm{KS}}$ as in~\eqref{eqn:kernel_smoothing}, using a kernel $\psi$ which satisfies~\eqref{asmp:boxcar_kernel}.
+Then there exist universal constants $c, N > 0$ such that for all $n > N$,
+		\begin{equation}
+		\label{eqn:cluster_assumption_ks_lb}
+		\inf_{h' > 0} \Ebb\Bigl[\|\wt{f} - f_0^{(n)}\|_{n}^2\Bigr] \geq c\min\biggl\{ \frac{r^{-1}}{n}, \frac{\theta}{\sqrt{n}} \biggr\}.
+		\end{equation}
+		\item Compute the least squares estimator $\wt{f} = \wt{f}_{\mathrm{SP}}$ as in~\eqref{eqn:least_squares}. Then there exist universal constants $c, N > 0$ such that for all $n > N$,
+		\begin{equation}
+		\label{eqn:cluster_assumption_sp_lb}
+		\inf_{1 \leq K \leq n} \Ebb \Bigl[\|\wt{f} - f_0^{(n)}\|_n^2\Bigr] \geq c\min\biggl\{\frac{r^{-1}}{n}, \frac{1}{\log(n)}, \frac{r^{-2/3}}{n}, \frac{\sqrt{\theta}}{n^{3/4}}\biggr\}.
 		\end{equation}
@@
 \end{proposition}
-The statements~\eqref{eqn:eigenmaps_beats_kernel_smoothing_1} and~\eqref{eqn:eigenmaps_beats_kernel_smoothing_2} are finite sample and hold for any $\theta$.
-Taking a sequence $(\theta_n)_{n \in \mathbb{N}}$ such that $\theta = o(\sqrt{n})$ and $\theta = \omega((\log n)^2/\sqrt{n})$ implies that
-\begin{equation*}
-\lim_{n \to \infty} \sup_{h'}  \frac{\Ebb \|T_{n,h}\wh{f} - f_0^{(n)}\|_{P^{(n)}}^2}{	\Ebb \|T_{n,h'}{\bf Y} - f_0^{(n)}\|_{P^{(n)}}^2 } = 0.
-\end{equation*}
-
-\paragraph{Proof of Proposition~\ref{prop:eigenmaps_beats_kernel_smoothing}.}
-First we prove~\eqref{eqn:eigenmaps_beats_kernel_smoothing_1}, then~\eqref{eqn:eigenmaps_beats_kernel_smoothing_2}.
-
-\underline{\emph{Proof of~\eqref{eqn:eigenmaps_beats_kernel_smoothing_1}}.}
-We begin by showing that with high probability the eigenvectors $v_1,v_2$ respect the cluster structure of $p$.
-Formally, letting $u_1 = (\1\{X_i \in Q_1\})_{i \in [n]}$, and likewise $u_2 = (\1\{X_i \in Q_2\})_{i \in [n]}$, we establish that
-\begin{equation}
-\label{pf:eigenmaps_beats_kernel_smoothing_1}
-\Pbb\Bigl(\mathrm{span}\{v_1,v_2\} = \mathrm{span}\{u_1,u_2\}\Bigr) \geq 1 - \frac{1}{n}.
-\end{equation}
-Note that~\eqref{pf:eigenmaps_beats_kernel_smoothing_1} is equivalent to the statement that the neighborhood graph $G_{n,\varepsilon}$ consists of exactly two connected components: one consisting of all design points $X_i \in Q_1$, and the other consisting of all design points $X_i \in Q_2$.
-Of course, since $h < r$ and $\eta$ is compactly supported on $[0,1]$, it follows that no $X_i \in Q_1$ and $X_j \in Q_2$ can be connected.
-On the other hand, using an elementary concentration argument (stated in Lemma~\ref{lem:balls_in_bins}) and the triangle inequality, we deduce that with probability at least $1 - 1/n$ there exists a path in $G_{n,\varepsilon}$ between each $X_i, X_j \in Q_1$, and likewise between each $X_i,X_j \in Q_2$.
-This establishes~\eqref{pf:eigenmaps_beats_kernel_smoothing_1}.
-
-Now we condition on the ``good'' event $\mc{E}$ that the design points $X_1,\ldots,X_n$ satisfy~\eqref{eqn:balls_in_bins}, and therefore that $\mathrm{span}\{v_1,v_2\} = \mathrm{span}\{u_1,u_2\}$.
-Consider the empirical mean $\wb{Y}_Q := \frac{1}{\sharp\{Q \cup {\bf X}\}} \sum_{i: X_i \in Q} Y_i$.
-Since $\mathrm{span}\{v_1,v_2\} = \mathrm{span}\{u_1,u_2\}$, the estimator $\wh{f}$ will be piecewise constant on $Q_1$ and $Q_2$, and in fact we have that
-\begin{equation*}
-\wh{f} = \wb{Y}_{Q_1}u_1 + \wb{Y}_{Q_2}u_2.
-\end{equation*}
-Moreover, by~\eqref{eqn:balls_in_bins} every $x \in Q_1$ is within distance $h$ of at least one $X_i \in {\bf X} \cup Q_1$ and likewise every $x \in Q_2$ is within distance $h$ of at least one $X_j \in {\bf X} \cup Q_2$.
-Consequently, the kernel smoothening of $\wh{f}$ also exhibits a piecewise constant structure,
-\begin{equation*}
-T_{n,h}\wh{f}(x) = \wb{Y}_{Q_1}\1\{x \in Q_1\} + \wb{Y}_{Q_2}\1\{x \in Q_2\},\quad \textrm{for all}~~x \in Q_1 \cup Q_2.
-\end{equation*}
-Letting
-\begin{equation*}
-\|T_{n,h}\wh{f} - f_0^{(n)}\|_{L^2(P^{(n)})}^2 = \frac{1}{2}\Bigl((\wb{Y}_{Q_1} - \theta)^2 + (\wb{Y}_{Q_2} + \theta)^2\Bigr),
-\end{equation*}
-it follows from Lemma~\ref{lem:chi_square_bound} that conditional on $\mc{E}$,
-\begin{equation}
-\label{pf:eigenmaps_beats_kernel_smoothing_0}
-\|T_{n,h}\wh{f} - f_0^{(n)}\|_{L^2(P^{(n)})}^2 \leq \frac{4(\log n)^2}{n}.
-\end{equation}
-
-Now we derive a crude upper bound on $\|T_{n,h}\wh{f} - f_0^{(n)}\|_{L^2(P^{(n)})}$ that will suffice to control the error conditional on $\mc{E}^c$. Note that if $d_{n,h}(x) = 0$ then
-$T_{n,h}\wh{f}(x) = 0$, and $\bigl(T_{n,h}\wh{f}(x) - f_0(x)\bigr)^2 = [f_0(x)]^2$. On the other hand, if $d_{n,h}(x) \geq 1$, then by Jensen's inequality it follows that
-\begin{align*}
-\bigl(T_{n,h}\wh{f}(x) - f_0(x)\bigr)^2 & \leq \biggl(\frac{1}{d_{n,h}(x)} \sum_{i = 1}^{n} (\wh{f}_i - f_0(x)) \cdot \1\{\|X_i - x\| \leq h \}\biggr)^2 \\
-& \leq \frac{1}{d_{n,h}(x)} \sum_{i = 1}^{n} (\wh{f}_i - f_0(x))^2 \cdot \1\{\|X_i - x\| \leq h \} \\
-& \leq 2[f_0(x)]^2 + \frac{2}{d_{n,h}(x)} \sum_{i = 1}^{n} (\wh{f}_i)^2 \cdot \1\{\|X_i - x\| \leq h\},
-\end{align*}
-and integrating over all $x \in \mc{X}^{(n)}$ gives
-\begin{align*}
-\|T_{n,h}\wh{f} - f_0\|_{\Leb^2(P^{(n)})}^2 & \leq 2 \|f_0\|_{\Leb^2(P^{(n)})}^2 + \sum_{i = 1}^{n} (\wh{f}_i)^2 \biggl(\int \frac{\1\{\|X_i - x\| \leq h \}}{d_{n,h}(x)} \,dP(x)\biggr) \\
-& \leq 2 \|f_0\|_{\Leb^2(P^{(n)})}^2 + 2h\sum_{i = 1}^{n} (\wh{f}_i)^2 \\
-& = 2 \|f_0\|_{\Leb^2(P^{(n)})}^2 + 2 (\log n)^2 \|\wh{f}\|_n^2.
-\end{align*}
-We can further upper bound the empirical norm of $\|\wh{f}\|_n^2$ using the Cauchy-Schwarz inequality:
-\begin{equation*}
-\|\wh{f}\|_n^2 \leq \frac{2}{n}\sum_{i = 1}^{n}\dotp{Y}{v_1}_n^2v_{1,i}^2 + \dotp{{\bf Y}}{v_2}_n^2v_{2,i}^2 \leq 2\bigl(\dotp{{\bf Y}}{v_1}_n^2 + \dotp{{\bf Y}}{v_2}_n^2\bigr) \leq 4 \|{\bf Y}\|_n^2.
-\end{equation*}
-Noting that $\Ebb[\|{\bf Y}\|_n^2|{\bf X}] = \|f_0\|_n^2 + 1/n$, and that $\max_x |f_0(x)| = \theta$, we conclude that
-\begin{equation*}
-\Ebb\Bigl[\|T_{n,h}\wh{f} - f_0\|_{\Leb^2(P^{(n)})}^2 \cdot \1\{\mc{E}^c\}\Bigr] \leq \Ebb\Bigl[\Bigl(2 \|f_0\|_{\Leb^2(P^{(n)})}^2 + 4 (\log n)^2 \|{\bf Y}\|_n^2\Bigr) \cdot \1\{\mc{E}^c\} \Bigr] \leq \frac{2 \theta^2}{n} + \frac{4 (\log n)^2(\theta^2 + 1)}{n}.
-\end{equation*}
-Combining this with~\eqref{pf:eigenmaps_beats_kernel_smoothing_0} implies~\eqref{eqn:eigenmaps_beats_kernel_smoothing_1}.
-
-\underline{\emph{Proof of~\eqref{eqn:eigenmaps_beats_kernel_smoothing_2}}.}
-Fix $x \in \mc{X}^{(n)}$. A standard argument using the law of iterated expectation implies the following lower bound on the pointwise risk in terms of squared-bias and variance-like quantities,
-\begin{equation*}
-\Ebb\Bigl[\Bigl(T_{n,h'}{\bf Y}(x) - f_0(x)\Bigr)^2\Bigr] \geq \Ebb\biggl[\Bigl(f_0(X) - f_0(x)\Bigr)^2|X \in B(x,h')\biggr] + \Ebb\biggl[\frac{1}{d_{n,h'}(x)}\biggr].
+The proof of Proposition~\ref{prop:cluster_assumption_lb} is long, and we defer it until after some discussion of the implications of the proposition.
+
+Together, Propositions~\ref{prop:cluster_assumption_ub} and~\ref{prop:cluster_assumption_lb} illustrate that the risk of PCR-LE can be dramatically smaller than that of kernel smoothing or uniform least squares.
+For instance, taking $\theta = n^{1/2}$ and $r = n^{-3/4}$, when appropriately tuned, $\wh{f}$ satisfies
+\begin{equation*}
+\Ebb\Bigl[\|\wh{f} - f_0^{(n)}\|_{n}^2\Bigr] \leq C\biggl(n^{7/4}\exp(-n^{1/4}/8)) + \frac{1}{n}\biggr) \leq \frac{C}{n},
+\end{equation*}
+for a universal constant $C$ and all $n$ larger than some universal constant $N$, whereas for $\wt{f} = \wt{f}_{\mathrm{KS}}$,
+\begin{equation*}
+\inf_{h' > 0} \Ebb\Bigl[\|\wt{f} - f_0^{(n)}\|_{n}^2\Bigr] \geq \frac{c}{n^{1/4}},
+\end{equation*}
+and for $\wt{f} = \wt{f}_{\mathrm{SP}}$,
+\begin{equation*}
+\inf_{1 \leq K \leq n} \Ebb\Bigl[\|\wt{f} - f_0^{(n)}\|_{n}^2\Bigr] \geq \frac{c}{n^{1/2}}.
+\end{equation*}
+Other choices of $\theta$ and $r$ lead to even more dramatic gaps between the risk of PCR-LE, and the risk of kernel smoothing and least squares.
+The overall takeaway is that under Model~\ref{def:model_cluster_assumption}, estimators that use the graph Laplacian can converge to the true regression function $f_0^{(n)}$ at fast rates---parametric rates that do not depend on the $L^2$ norm of $f_0^{(n)}$---whereas other estimators, optimal for estimation over Sobolev spaces, converge to $f_0^{(n)}$ at slow rates---nonparametric rates that deteriorate as the $L^2$ norm of $f_0^{(n)}$ grows.
+
+Some remarks:
+\begin{itemize}
+	\item The lower bound on the in-sample risk of $\wt{f}_{\mathrm{KS}}$  given by \eqref{eqn:cluster_assumption_ks_lb} is larger than that of $\wt{f}_{\mathrm{SP}}$ given by \eqref{eqn:cluster_assumption_sp_lb}.
+This does not mean that kernel smoothing exhibits less adaptivity to the cluster assumption than uniform least squares.
+Instead, we suspect it is due to looseness in our lower bounds: we are able to tightly control the bias of kernel smoothing, whereas we must use a potentially loose bound on the bias of uniform least squares.
+Experimentally, it appears that kernel smoothing usually outperforms uniform least squares, under various instantiations of the cluster assumption.
+	\item The cluster assumption---in which the regression function is piecewise constant and $p$ consists of multiple connected components---is a very strong assumption.
+The \emph{low-density separation} condition is a related but weaker assumption, in which the regression function is assumed to be smoother (but not constant) in regions of higher density.
+This is a rather general hypothesis which can formalized in a number of different ways.
+For instance, one could insist that the regression function $f_0$ belong to a normed ball in a \emph{weighted Sobolev space}, with semi-norm given by
+	\begin{equation*}
+	|f_0|_{H^s(P)} := \dotp{\Delta_P^s f_0}{f_0}_P.
+	\end{equation*}
+	Intuitively, when $\|f_0\|_{H^s(P)}$ is much smaller than $\|f_0\|_{H^s(\mc{X})}$, density-adaptive learners such as PCR-LE should have the advantage on non-density adaptive linear smoothers, such as kernel smoothing or uniform least squares.
+Indeed, in the case of Model~\ref{def:model_cluster_assumption} we see that
+	\begin{equation*}
+	\|f_0^{(n)}\|_{H^s(P^{(n)})} = 0~~\textrm{for all $s \in \mathbb{N}$, and all $r, \theta > 0$,}
+	\end{equation*}
+	whereas $f_0^{(n)}$ does not even belong to the first-order Sobolev space $H^1([0,1])$.
+In words, this shows the cluster assumption is an extreme case of the low-density separation condition.
+Unfortunately, it is quite difficult to analyze graph-based estimators under the general low-density separation condition, without making strong assumptions on $P$.
+	\item Finally, we note that either changing the graph or the normalization of the Laplacian fundamentally alters the type of density adaptivity displayed by graph-Laplacian-based estimators; see~\citet{hoffmann2019} for an extensive discussion.
+\end{itemize}
+
+\subsection{Proof of Proposition~\ref{prop:cluster_assumption_lb}}
+First we show~\eqref{eqn:cluster_assumption_ks_lb}, then~\eqref{eqn:cluster_assumption_sp_lb}.
+
+\subsubsection{Proof of~\eqref{eqn:cluster_assumption_ks_lb}}
+A standard argument using the law of iterated expectation implies the following lower bound on the pointwise risk in terms of squared-bias and variance-like quantities,
+\begin{equation*}
+\Ebb\Bigl[\Bigl(\wt{f}(X_i) - f_0(X_i)\Bigr)^2|X_i = x\Bigr] \geq \frac{(n - 1)}{n}\Ebb\biggl[\Bigl(f_0(X) - f_0(x)\Bigr)^2|X \in B(x,h')\biggr] + \Ebb\biggl[\frac{1}{d_{n,h'}(x)}\biggr].
 \end{equation*}
@@
 \end{equation*}
-Combining these lower bounds on variance and squared bias terms and integrating over $x' \in \mc{X}$, we deduce the following: if $h' \leq 4r$, then
-\begin{equation*}
-\Ebb\Bigl[\|T_{n,h'}{\bf Y} - f_0^{(n)}\|_{L^2(P^{(n)})}^2\Bigr] \geq \frac{1}{16(\log n)^2},
+Combining these lower bounds on variance and squared bias terms and summing over $X_1,\ldots,X_n$, we arrive at the following: if $h' \leq 4r$, then
+\begin{equation*}
+\Ebb\Bigl[\|\wt{f} - f_0^{(n)}\|_{n}^2\Bigr] = \frac{1}{n}\sum_{i = 1}^{n} \Ebb\biggl[\Ebb\Bigl[\Bigl(\wt{f}(X_i) - f_0(X_i)\Bigr)^2|X_i\Bigr]\biggr]\geq \frac{1}{16rn},
 \end{equation*}
 whereas if $h' > 4r$  then
-\begin{equation*}\Ebb\Bigl[\|T_{n,h'}{\bf Y} - f_0^{(n)}\|_{L^2(P^{(n)})}^2\Bigr] \geq \frac{1}{4nh'} + \frac{\theta^2h'}{32}.
-\end{equation*}
-In the latter case, setting the derivative equal to $0$ shows that the right hand side is always at least $\theta/\sqrt{32 n}$, and taking the minimum over the two cases then yields~\eqref{eqn:eigenmaps_beats_kernel_smoothing_2}.
-
-\paragraph{A Useful Lemma.}
-Let $m = 4n/\log^2n$. Furthermore, for $i = 0,\ldots,m - 1$, let
-\begin{equation*}
-Q_{i1} = [i/m,(i + 1)/m] \cdot (1/2 - r), \quad Q_{i2} = 1/2 + [i/m,(i + 1)/m] \cdot (1/2 - r).
-\end{equation*}
-\begin{lemma}
-	\label{lem:balls_in_bins}
-	Suppose $(X_1,Y_1),\ldots(X_n,Y_n)$ are sampled according to~\eqref{def:model_cluster_assumption}. For any $\log(n) > 16$, we have that 
-	\begin{equation}
-	\label{eqn:balls_in_bins}
-	\Pbb\Bigl(\sharp\{Q_{ij} \cup {\bf X}\} > 0 ~~\textrm{for all $i = 1,\ldots,m - 1$ and $j = 1,2$} \Bigr) \geq 1 - \frac{1}{n}.
-	\end{equation}
-\end{lemma}
-Let $\varepsilon = h = \log^2n/2n$.
-Note that by construction,~\eqref{eqn:balls_in_bins} implies that any points $x$ and $x'$ in adjacent intervals $Q_{ij}$ and $Q_{i'j}$ must be connected in $G_{n,\varepsilon}$.
-It also implies that $d_{n,h}(x) > 0$ for every $x \in Q_1 \cup Q_2$.
-
-\paragraph{Proof (of Lemma~\ref{lem:balls_in_bins}).}
-For each $Q_{ij}$, we have that 
-\begin{equation*}
-\Pbb\bigl(\sharp\{Q_{ij} \cup {\bf X}\} = 0\bigr) = (1 - 1/(2m))^{n} \leq \exp\{-n/2m\}.
-\end{equation*}
-Thus by a union bound,
-\begin{equation*}
-\Pbb\Bigl(\sharp\{Q_{ij} \cup {\bf X}\} = 0 ~~\textrm{for any $i = 1,\ldots,m - 1$ and $j = 1,2$} \Bigr) \leq 2m\exp\{-n/2m\} = \frac{8n}{\log^2n}\exp\{-\log^2n/8\},
-\end{equation*}
-and the claim follows from some basic algebra.
+\begin{align*}
+\Ebb\Bigl[\|\wt{f} - f_0^{(n)}\|_{n}^2\Bigr] & = \frac{1}{n}\sum_{i = 1}^{n} \Ebb\biggl[\Ebb\Bigl[\Bigl(\wt{f}(X_i) - f_0(X_i)\Bigr)^2|X_i\Bigr]\biggr] \\
+& \geq \frac{1}{4nh'} + \frac{\theta^2}{8}\frac{(n - 1)}{n} P^{(n)}\Bigl([(1 - h'/2)_{+}, 1/2 - r]\Bigr) \\
+&\geq \frac{1}{4nh'} + \frac{\theta^2h'}{64}.
+\end{align*}
+In the latter case, setting the derivative equal to $0$ shows that the right hand side is always at least $\theta/\sqrt{64 n}$, and taking the minimum over the two cases then yields~\eqref{eqn:cluster_assumption_ks_lb}.
+
+\subsubsection{Proof of~\eqref{eqn:cluster_assumption_sp_lb}}
+We begin by decomposing the risk into conditional bias and variance terms.
+Let $\mathbb{E}_n = \Ebb[\cdot|X_1,\ldots,X_n]$ denote expectation conditional on the design points $X_1,\ldots,X_n$.
+Then by the law of iterated expectation, and the fact that $\Ebb_n[w] = 0$,
+\begin{equation*}
+\Ebb\bigl[\|\wt{f}_{\mathrm{SP}} - f_0\|_n^2\bigr] = \Ebb\bigl[\|\Ebb_n\wt{f}_{\mathrm{SP}} -  f_0\|_n^2\bigr] + \Ebb\bigl[\|\wt{f}_{\mathrm{SP}} - \Ebb_n\wt{f}_{\mathrm{SP}}\|_n^2\bigr]. 
+\end{equation*}
+We separately lower bound the expected conditional squared bias and variance terms.
+To anticipate what is to come: we will show that the expected conditional variance is equal to $K/n$; on the other hand we will show that the expected conditional squared bias is lower bounded,
+\begin{equation}
+\label{pf:cluster_assumption_sp_lb_1}
+\Ebb\bigl[\|\Ebb_n\wt{f}_{\mathrm{SP}} -  f_0\|_n^2\bigr] = \frac{K}{n}\quad\textrm{and}\quad\Ebb\bigl[\|\Ebb_n\wt{f}_{\mathrm{SP}} -  f_0\|_n^2\bigr] \geq \frac{\theta^2}{2601\pi^2 K^{3}},
+\end{equation}
+with the lower bound holding so long as $K \leq \min\{1/(16r),n/(8 \log(8 n)), (\sqrt{160} \pi/r)^{2/3}\}$. If $K$ is larger than this, then the expected conditional variance is lower bounded,
+\begin{equation}
+\label{pf:cluster_assumption_sp_lb_2}
+\Ebb\bigl[\|\wt{f}_{\mathrm{SP}} - \Ebb_n\wt{f}_{\mathrm{SP}}\|_n^2\bigr] \geq \min\biggl\{\frac{1}{16rn}, \frac{1}{8 \log(8n)}, \frac{(\sqrt{160} \pi)^{2/3}}{r^{2/3}n} \biggr\}
+\end{equation}
+Otherwise~\eqref{pf:cluster_assumption_sp_lb_1} implies that the in-sample risk is always at least 
+\begin{equation*}
+\Ebb\bigl[\|\wt{f}_{\mathrm{SP}} - f_0\|_n^2\bigr] \geq \frac{\theta^2}{2601\pi^2 K^{3}} + \frac{K}{n} \geq 2 \frac{\theta^{1/2}}{n^{3/4}} \frac{1}{(2601 \pi^2)^{1/4}}.
+\end{equation*}
+Along with~\eqref{pf:cluster_assumption_sp_lb_2}, this implies the claim. It remains to show the bounds on conditional bias and variance.
+
+\paragraph{Conditional variance.}
+The expected conditional variance is exactly equal to $K/n$, a standard fact that is verified by the following calculations: first,
+\begin{align*}
+\|\wt{f}_{\mathrm{SP}} - \Ebb_n\wt{f}_{\mathrm{SP}}\|_n^2 = \|\Phi(\Phi^{\top}\Phi)^{-1}\Phi^{\top}w\|_n^2 = \frac{1}{n}w^{\top}\Phi(\Phi^{\top}\Phi)^{-1}\Phi^{\top}w;
+\end{align*}
+thus standard properties of the Gaussian distribution and the trace trick imply
+\begin{equation*}
+\Ebb_n\bigl[\|\wt{f}_{\mathrm{SP}} - \Ebb_n\wt{f}_{\mathrm{SP}}\|_n^2 \bigr] = \frac{1}{n} \mathrm{tr} (\Phi(\Phi^{\top}\Phi)^{-1}\Phi^{\top}) = \frac{K}{n};
+\end{equation*}
+and finally by the law of iterated expectation and the independence of the noise $(w_1,\ldots,w_n)$ and the design points $X_1,\ldots,X_n$,
+\begin{equation*}
+\Ebb\Bigl[\Ebb_n\bigl[\|\wt{f}_{\mathrm{SP}} - \Ebb_n\wt{f}_{\mathrm{SP}}\|_n^2\Bigr] = K/n.
+\end{equation*}
+
+\paragraph{Conditional bias.}
+It takes more work to lower bound the conditional bias.
+We will first upper bound the Lipschitz constant of $\Ebb_n\wt{f}_{\mathrm{SP}}$ in terms of the empirical norm $\|\Ebb_n\wt{f}_{\mathrm{SP}}\|_n$.
+Then we will use this upper bound to argue that either $\Ebb_n\wt{f}_{\mathrm{SP}}$ has empirical norm much larger than that of $f_0$, or $\Ebb_n\wt{f}_{\mathrm{SP}}$ is a smooth function, in the sense of having a small Lipschitz constant.
+In the former case, the triangle inequality will then imply that $\|\Ebb_n\wt{f}_{\mathrm{SP}} -  f_0\|_n$ must be large.
+In the latter case, the smoothness of $\Ebb_n\wt{f}_{\mathrm{SP}}$ will imply that $\Ebb_n\wt{f}_{\mathrm{SP}}$ must be far from $f_0$ at many points $X_i$ close to $x = 1/2$.
+
+The following Lemma gives our upper bound on the Lipschitz constant of $\|\Ebb_n\wt{f}_{\mathrm{SP}}\|_n$.
+Here we treat $\Ebb_n\wt{f}_{\mathrm{SP}} = \sum_{k = 1}^{K} \wt{\beta}_k \phi_k$ as a function defined at all $x \in [0,1]$ by extending it in the canonical way.
+As a function over $[0,1]$, clearly $\Ebb_n\wt{f}_{\mathrm{SP}} \in C^\infty([0,1])$.
+Let $\Sigma \in \Reals^{K \times K}$ be the covariance matrix of $(\phi_1,\ldots,\phi_K)$, i.e.
+the matrix with entries $\Sigma_{k\ell} = \dotp{\phi_k,\phi_{\ell}}_{P^{(n)}}$.
+Let $\wh{\Sigma} := (\Phi^{\top}\Phi)/n$ be the empirical covariance matrix.
+Let $I_K \in \Reals^{K \times K}$ be the identity matrix.
+\begin{lemma}[Lipschitz regularity of $\Ebb_n\wt{f}_{\mathrm{SP}}$.]
+	\label{lem:lipschitz_regularity_sp}
+	Let $\wt{f}_n = \Ebb_n\wt{f}_{\mathrm{SP}}$. Then 
+	\begin{equation}
+	\label{eqn:lipschitz_regularity_sp1}
+	\|\wt{f}_n\|_{C^1(\mc{X})}^2 \leq \pi^2 \frac{K^{3} \cdot \|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\|_{\mathrm{op}}}{(1 - \|I_K - \Sigma\|_F)} \cdot \|\wt{f}_n\|_n^2.
+	\end{equation}
+	Moreover, suppose $K \leq 1/(16r)$ and $r \leq (1 - 2^{-1/2})/2$.
+	\begin{itemize}
+		\item {\bf(Matrix perturbation)} Then
+		\begin{equation}
+		\label{eqn:lipschitz_regularity_matrix_perturbation}
+		\|\Sigma - I_K\|_F \leq \frac{1}{2}.
+		\end{equation}
+		\item {\bf(Matrix concentration, cf. \cite{hsu2012})} If additionally $n \geq 8K \log(K/\delta)$ for some $\delta \in (0,1/2)$, then with probability at least $1 - 2\delta$,
+		\begin{equation}
+		\label{eqn:lipschitz_regularity_matrix_concentration}
+		\|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\|_{\mathrm{op}} \leq 5.
+		\end{equation}
+	\end{itemize}
+	Therefore, if $K \leq \min\{1/(16r),n/(8 \log(K/\delta))\}$, then with probability at least $1 - 2\delta$,
+	\begin{equation}
+	\label{eqn:lipschitz_regularity_sp2}
+	\|\wt{f}_n\|_{C^1(\mc{X})}^2 \leq 10 \pi^2 K^3 \|\wt{f}_n\|_n^2.
+	\end{equation}
+\end{lemma}
+We defer the proof of Lemma~\ref{lem:lipschitz_regularity_sp} until after we complete the proof of~\eqref{eqn:cluster_assumption_sp_lb}.
+
+Now, if $\|\wt{f}_n\|_n^2 \geq \frac{3}{2}\|f_0\|_n^2$, then by the triangle inequality
+\begin{equation*}
+\|\wt{f}_n - f_0\|_n \geq \|\wt{f}_n\|_n - \|f_0\|_n \geq \sqrt{\frac{3}{2}} \cdot \|f_0\|_n  = \sqrt{\frac{3}{2}} \cdot \theta.
+\end{equation*}
+Otherwise $\|\wt{f}_n\|_n^2 \geq \frac{3}{2}\|f_0\|_n^2$.
+In this case, we show that $|\wt{f}_n(X_i) - f_0(X_i)|$ must be large (on the order of $\theta$) for many points $X_i$ which are close to $x = 1/2$.
+Let us suppose without loss of generality that $\wt{f}_n(1/2) \leq \theta/2$ and consider points $X_i \in Q_1$ close to $x = 1/2$; otherwise if $\wt{f}_n(1/2) > \theta/2$ we could obtain the exact same bound by considering $X_i \in Q_2$.
+For each point $X_i \in Q_1$, by Lemma~\ref{lem:lipschitz_regularity_sp} we have that with probability at least $1 - 2\delta$,
+\begin{equation*}
+|\wt{f}_n(X_i) - \wt{f}_n(1/2)| \leq CK^{3/2} \|\wt{f}_n\|_n \cdot |X_i - 1/2| \leq \sqrt{10}\pi K^{3/2} \theta \cdot |X_i - 1/2|.
+\end{equation*}
+Since $\wt{f}_n(1/2) \leq \theta/2$ and $f_0(X_i) = \theta/2$ for all $X_i \in Q_1$ it follows that
+\begin{equation*}
+|\wt{f}_n(X_i) - f_0(X_i)| \geq \theta - \sqrt{10}\pi K^{3/2} \theta \cdot |X_i - 1/2|,
+\end{equation*}
+and consequently
+\begin{equation*}
+|\wt{f}_n(X_i) - f_0(X_i)| \geq \theta/2,\quad\textrm{for any $X_i \in Q_1$ such that $|X_i - 1/2| \leq 1/(\sqrt{40}\pi K^{3/2})$.}
+\end{equation*}
+This yields a lower bound on $\|\wt{f}_n - f_0\|_n$; letting $Q_K := \Bigl[\frac{1}{2} - \frac{1}{\sqrt{40}\pi K^{3/2}},\frac{1}{2} - r\Bigr]$, we have that 
+\begin{equation*}
+\|\wt{f}_n - f_0\|_n \geq \frac{\theta}{2} \cdot P_n\Bigl(Q_k\Bigr).
+\end{equation*}
+Then as long as $K^{-3/2} \geq \sqrt{160} \pi r$, from the multiplicative form of Hoeffding's inequality (Lemma~\ref{lem:hoeffding_smoothing})
+\begin{equation*}
+P^{(n)}(Q_K) \geq \frac{1}{\sqrt{160} \pi K^{3/2}} \geq 2r \Longrightarrow \Pbb\biggl(P_n(Q_K) \geq \frac{1}{\sqrt{640} \pi K^{3/2}}\biggr) \geq 1 - \exp(-nr/4) \geq 1 - \frac{4}{n^2}.
+\end{equation*}
+Putting the pieces together, we conclude that if $K \leq \min\{1/(8r),n/(8 \log(K/\delta)), (\sqrt{160} \pi/r)^{2/3}\}$, then
+\begin{equation*}
+\|\wt{f}_n - f_0\|_n \geq \frac{\theta}{51 \pi K^{3/2}},
+\end{equation*}
+with probability at least $1 - 2\delta - {4}{n^2}$. Taking $\delta = 1/8$ then implies the claim.
+
+\paragraph{Proof of Lemma~\ref{lem:lipschitz_regularity_sp}.}
+
+\underline{\it{Proof of~\eqref{eqn:lipschitz_regularity_sp1}}.}
+Recall that $\wt{f}_n = \sum_{k = 1}^{K} \wt{\beta}_k \phi_k$. Exchanging sum with derivative, we have that
+\begin{equation*}
+\frac{d}{dx} \wt{f}_n(x) = -\pi \sum_{k = 1}^{K}\wt{\beta}_k k \sin(k\pi x).
+\end{equation*}
+Thus taking absolute value and applying the Cauchy-Schwarz inequality gives
+\begin{equation*}
+|\wt{f}_n'(x)|^2 \leq \pi^2 K^2 \sum_{k = 1}^{K} \bigl(\sin(k\pi x)\bigr)^2 \|\beta\|_2^2 \leq \pi^2 K^3 \|\beta\|_2^2.
+\end{equation*}
+On the other hand, we can also relate the empirical norm $\|\wt{f}_n\|_n^2$ to the $\ell^2$ norm of $\beta$. Specifically,
+\begin{equation*}
+\|\wt{f}_n\|_n^2 = \beta^{\top} \wh{\Sigma} \beta \geq \frac{\|\beta\|_2^2}{\|\wh{\Sigma}^{-1}\|_{\mathrm{op}}} \geq \frac{\|\beta\|_2^2}{\|\Sigma^{-1}\|_{\mathrm{op}} \cdot \|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\|_{\mathrm{op}}} = \frac{\|\beta\|_2^2 \|\Sigma\|_{\mathrm{op}} }{\|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\|_{\mathrm{op}}}
+\end{equation*}
+Rearranging, we see that 
+\begin{equation*}
+\sup_{x \in [0,1]} |\wt{f}_n'(x)|^2 \leq \frac{\pi^2 K^3}{\|\Sigma\|_{\mathrm{op}}} \|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\|_{\mathrm{op}} \leq \frac{\pi^2 K^3}{1 - \|I_K - \Sigma\|_F} \|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\|_{\mathrm{op}}
+\end{equation*}
+with the latter inequality following since $\|\Sigma\|_{\mathrm{op}} \geq \|I_K\|_{\mathrm{op}} - \|I_K - \Sigma\|_{\mathrm{op}} \geq 1 - \|I_K - \Sigma\|_{F}$.
+
+\underline{\it{Proof of~\eqref{eqn:lipschitz_regularity_matrix_perturbation}}.}
+We will show that for all $1 \leq k < \ell \leq K$, 
+\begin{equation}
+\label{pf:lipshcitz_regularity_matrix_perturbation_1}
+(1 - \dotp{\phi_k}{\phi_k}_{P^{(n)}})^2 \leq 32r^2, \quad\textrm{and}\quad|\dotp{\phi_k}{\phi_{\ell}}_{P^{(n)}}| \leq 64r^2.
+\end{equation}
+This implies $\|I - \Sigma\|_F^2 \leq 32K^2r^2$, so that $\|I - \Sigma\|_F \leq 1/2$ so long as $K \leq 1/(16r)$.
+
+The proof of~\eqref{pf:lipshcitz_regularity_matrix_perturbation_1} follows from computing some standard integrals. We separate the computation based on whether $k = 1$ or $k > 1$. \newline
+
+\underline{\it{Case 1: $k = 1$.}}
+When $k = 1$, $\dotp{\phi_1}{\phi_1}_{P^{(n)}} = 1$ and $(1 - \dotp{\phi_1}{\phi_1}_{P^{(n)}})^2 = 0$. Additionally, by symbolic integration we find that
+\begin{equation*}
+\dotp{\phi_k}{\phi_{\ell}}_{P^{(n)}} = \frac{-2\sqrt{2}}{(1 - 2r)} \cdot \frac{\cos(\ell\pi/2) \sin(\ell\pi r)}{\ell \pi},
+\end{equation*}
+and therefore
+\begin{equation*}
+\bigl[\dotp{\phi_k}{\phi_{\ell}}_{P^{(n)}}\bigr]^2 \leq \frac{8}{(1 - 2r)^2} \cdot \biggl(\frac{\sin(\ell\pi r)}{\ell \pi}\biggr)^2 \leq \frac{8}{(1 - 2r)^2} r^2 \leq 16r^2,
+\end{equation*}
+where in the second-to-last inequality follows because $\sin(x)/x \leq 1$, and the last inequality follows by our assumed upper bound on $r$. \newline
+
+\underline{\it{Case 2: $k > 1$.}}
+When $k > 1$,
+\begin{equation*}
+\dotp{\phi_k}{\phi_k}_{P^{(n)}} = 1 - \frac{2}{(1 - 2r)} \frac{\cos(k\pi) \sin(2k\pi r)}{k\pi} \Longrightarrow \bigl[1 - \dotp{\phi_k}{\phi_k}_{P^{(n)}}\bigr]^2 \leq \frac{4}{(1 - 2r)^2} \cdot \biggl(\frac{\sin(2k\pi r)}{k\pi}\biggr)^2 \leq 32r^2.
+\end{equation*}
+Similarly,
+\begin{equation*}
+\dotp{\phi_k}{\phi_\ell}_{P^{(n)}} = -\frac{4}{(1 - 2r)}\biggl[\frac{\cos((k + \ell)\pi) \sin((k + \ell)\pi r)}{(k + \ell)\pi} + \frac{\cos((k - \ell)\pi) \sin((k - \ell)\pi r)}{(k - \ell)\pi}\biggr]
+\end{equation*}
+and therefore
+\begin{equation*}
+\bigl[\dotp{\phi_k}{\phi_\ell}_{P^{(n)}}\bigr]^2 \leq \frac{16}{(1 - 2r)^2} \biggl(\biggl[\frac{\sin((k + \ell)\pi r)}{(k + \ell)\pi}\biggr]^2 + \biggl[\frac{\sin((k - \ell)\pi r)}{(k - \ell)\pi}\biggr]^2\biggr) \leq 64 r^2.
+\end{equation*}
+
+\underline{\it{Proof of~\eqref{eqn:lipschitz_regularity_matrix_concentration}}}
+Denote $\Phi(x) = (\phi_1,\ldots,\phi_K(x)) \in \Reals^K$ for any $x \in [0,1]$. Then for any $x \in [0,1]$,
+\begin{equation*}
+\|\Sigma^{-1/2} \Phi(x)\| \leq \|\Sigma^{-1}\|_{\mathrm{op}}^{1/2} \|\Phi(x)\|_2 \leq \|\Sigma^{-1}\|_{\mathrm{op}}^{1/2} \sqrt{2K} \leq 2 \sqrt{K}
+\end{equation*}
+with the second-to-last inequality following from~\eqref{eqn:lipschitz_regularity_matrix_perturbation}, and the last inequality following since $|\phi_k(x)| \leq \sqrt{2}$ for all $k$.
+Thus $\|\Sigma^{-1/2} \Phi(x)\|/\sqrt{K} \leq 2$, and~\eqref{eqn:lipschitz_regularity_matrix_concentration} follows from Theorem~1 of \citet{hsu2012}.
+
+\underline{\it{Proof of~\eqref{eqn:lipschitz_regularity_sp2}}.}
+Follows immediately.
 
@@
 
-In the proof of Lemma~\ref{lem:kernel_smoothing_insample}, we require uniform control of the empirical degree functional $d_{n,h}(x)$ over all $x \in \mc{X}$.
-Such a result is available to us because the kernel $\psi$ is Lipschitz on its support, so that the class of functions $\{\psi((x - \cdot)/h): h \in \Rd\}$ has finite VC dimension.
-The precise estimate we use is due to~\cite{gine2002}.
-\begin{lemma}[Uniform bound for empirical degree.]
-	\label{lem:uniform_bound_empirical_degree}
-	Suppose Model~\ref{def:model_flat_euclidean}.
-For a kernel $\psi$ satisfying~\ref{asmp:kernel} and bandwidth $h$ satisfying~\ref{asmp:bandwidth}, there exist constants $c, C, c_1$ and $C_1$ which do not depend on $h$ or $n$ such that
-	\begin{equation*}
-	\Pbb\biggl(\sup_{x \in \Rd} n \cdot \Bigl|d_{n,h}(x) - d_{P,h}(x)\Bigr| > t\biggr) \leq C\exp\biggl(-c \frac{t^2}{nh^d}\biggr),
-	\end{equation*}
-	for any $t \in \Reals$ satisfying
-	\begin{equation}
-	\label{eqn:uniform_bound_empirical_degree_1}
-	C_1\sqrt{nh^d \log(1/h)} \leq t \leq c_1 nh^d.
-	\end{equation}
-\end{lemma}
-Now we translate Lemma~\ref{lem:uniform_bound_empirical_degree} into a multiplicative bound, which will be more useful for our purposes. 
-Recall the lower bound on $d_{P,h}(x)$ given in~\eqref{eqn:degree_lower_bound}.
-By setting $C_0$ to be a sufficiently large constant in~\ref{asmp:bandwidth}, we can ensure that choosing $t = n h^d p_{\min}/6$ satisfies both the inequalities~\eqref{eqn:uniform_bound_empirical_degree_1}.
-For this choice of $t$ it follows from~\eqref{eqn:degree_lower_bound} and Lemma~\ref{lem:uniform_bound_empirical_degree} that
-\begin{equation}
-\label{eqn:uniform_bound_empirical_degree_2}
-\sup_{x \in \Rd} \biggl|\frac{d_{n,h}(x) - d_{P,h}(x)}{d_{P,h}(x)}\biggr| \leq \sup_{x \in \Rd} \biggl|\frac{d_{n,h}(x) - d_{P,h}(x)}{2t/n}\biggr| \leq \frac{1}{2}
-\end{equation}
-with probability at least $1 - C\exp(-cnh^d)$. This is the form of the result we use in the proof of Lemma~\ref{lem:kernel_smoothing_insample}.
+Lemma~\ref{lem:hoeffding_smoothing} is a multiplicative form of Hoeffding's inequality.
+\begin{lemma}[Hoeffding's Inequality, multiplicative form]
+	\label{lem:hoeffding_smoothing}
+	Suppose $Z_i$ are independent random variables, which satisfy $Z_i \in [0,B]$ for $i = 1,\ldots,n$. For any $0 < \delta < 1$, it holds that
+	\begin{equation*}
+	\Pbb\biggl(\Bigl|S_n - \mu\Bigr| \geq \delta \mu\biggr) \leq 2\exp\biggl(-\frac{\delta^2\mu}{3B^2}\biggr).
+	\end{equation*}
+\end{lemma}
+
+The following Lemma gives a ``balls-in-bins'' result. More precisely, it gives a lower bound on the probability that every bin
+\begin{equation*}
+Q_{i1} = [i/m,(i + 1)/m] \cdot (1/2 - r), \quad Q_{i2} = 1/2 + [i/m,(i + 1)/m] \cdot (1/2 - r).
+\end{equation*}
+will contain at least one ball.
+\begin{lemma}
+	\label{lem:balls_in_bins}
+	Suppose $(X_1,Y_1),\ldots(X_n,Y_n)$ are sampled according to~\eqref{def:model_cluster_assumption}, and suppose $r \leq 1/4$. We have that 
+	\begin{equation}
+	\label{eqn:balls_in_bins}
+	\Pbb\Bigl(\sharp\{Q_{ij} \cup {\bf X}\} > 0 ~~\textrm{for all $i = 1,\ldots,m - 1$ and $j = 1,2$} \Bigr) \geq 1 - 2m\exp\{-n/2m\}.
+	\end{equation}
+\end{lemma}
+\paragraph{Proof (of Lemma~\ref{lem:balls_in_bins}).}
+For each $Q_{ij}$, we have that $P(Q_{ij}) = (1/2 - r)/m \geq 1/(2m)$. Therefore
+\begin{equation*}
+\Pbb\bigl(\sharp\{Q_{ij} \cup {\bf X}\} = 0\bigr) = (1 - 1/(2m))^{n} \leq \exp\{-n/2m\}.
+\end{equation*}
+By a union bound,
+\begin{equation*}
+\Pbb\Bigl(\sharp\{Q_{ij} \cup {\bf X}\} = 0 ~~\textrm{for any $i = 1,\ldots,m - 1$ and $j = 1,2$} \Bigr) \leq 2m\exp\{-n/2m\}. \qed
+\end{equation*}
+Let $\varepsilon = 2/m$.
+Note that by construction,~\eqref{eqn:balls_in_bins} implies that any points $x$ and $x'$ in adjacent intervals $Q_{ij}$ and $Q_{i'j}$ must be connected in $G_{n,\varepsilon}$.
+Likewise, it implies that for $h = 1/m$ the degree $d_{n,h}(x) > 0$ for every $x \in Q_1 \cup Q_2$.
 
@@
 \label{sec:computational_considerations}
-Recall that when $s = 1$, we have shown that Laplacian eigenmaps is optimal when $\varepsilon \asymp (\log n/n)^{1/d}$ is (up to a constant) as small as possible while still ensuring the graph $G$ is connected.
-On the other hand, when $s > 1$, we can show Laplacian eigenmaps is optimal only when $\varepsilon = \omega(n^{-c})$ for some $c < 1/d$.
+Recall that when $s = 1$, we have shown that PCR-LE is optimal when $\varepsilon \asymp (\log n/n)^{1/d}$ is (up to a constant) as small as possible while still ensuring the graph $G$ is connected.
+On the other hand, when $s > 1$, we can show PCR-LE is optimal only when $\varepsilon = \omega(n^{-c})$ for some $c < 1/d$.
 For such a choice of $\varepsilon$, the average degree in $G$ will grow polynomially in $n$ as $n \to \infty$, and computing eigenvectors of the Laplacian of a graph will be more computationally intensive than if the graph were sparse.
-In this dense-graph setting, we now discuss a procedure to more efficiently compute an approximation to the Laplacian eigenmaps estimate: \emph{edge sparsification}.
+In this dense-graph setting, we now discuss a procedure to more efficiently compute an approximation to the PCR-LE estimate: \emph{edge sparsification}.
 
@@
 \end{equation*}
-Let $\wc{f}$ be the Laplacian eigenmaps estimator computed using the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ .
+Let $\wc{f}$ be the PCR-LE estimator computed using the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ .
 Because $\wc{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wc{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$, and consequently much faster to compute $\wc{f}$ than $\wh{f}$.

### 2021-10-16 manuscripts/laplacian_eigenmaps/arxiv/introduction.tex
Updated Laplacian Eigenmaps manuscript.

\documentclass{article}

%%% Begin Ryan's template
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{booktabs}       % professional-quality tables
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{times}          % times font

\usepackage[round]{natbib}
\usepackage{amssymb,amsmath,amsthm,bbm}
\usepackage[margin=1in]{geometry}
\usepackage{verbatim,float,url,dsfont}
\usepackage{graphicx,subfigure,psfrag}
\usepackage{algorithm,algorithmic}
\usepackage{mathtools,enumitem}
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}
\usepackage{multirow}

% Theorems and such
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{remark}{Remark}
\newtheorem{definition}{Definition}

% Assumption
\newtheorem*{assumption*}{\assumptionnumber}
\providecommand{\assumptionnumber}{}
\makeatletter
\newenvironment{assumption}[2]{
	\renewcommand{\assumptionnumber}{Assumption #1#2}
	\begin{assumption*}
		\protected@edef\@currentlabel{#1#2}}
	{\end{assumption*}}
\makeatother

% Widebar
\makeatletter
\newcommand*\rel@kern[1]{\kern#1\dimexpr\macc@kerna}
\newcommand*\widebar[1]{%
	\begingroup
	\def\mathaccent##1##2{%
		\rel@kern{0.8}%
		\overline{\rel@kern{-0.8}\macc@nucleus\rel@kern{0.2}}%
		\rel@kern{-0.2}%
	}%
	\macc@depth\@ne
	\let\math@bgroup\@empty \let\math@egroup\macc@set@skewchar
	\mathsurround\z@ \frozen@everymath{\mathgroup\macc@group\relax}%
	\macc@set@skewchar\relax
	\let\mathaccentV\macc@nested@a
	\macc@nested@a\relax111{#1}%
	\endgroup
}
\makeatother

% Min and max
\newcommand{\argmin}{\mathop{\mathrm{argmin}}}
\newcommand{\argmax}{\mathop{\mathrm{argmax}}}
\newcommand{\minimize}{\mathop{\mathrm{minimize}}}
\newcommand{\st}{\mathop{\mathrm{subject\,\,to}}}

% Shortcuts
\def\R{\mathbb{R}}

%%% End Ryan's template

%%% Begin Alden's additions
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\Reals}{\mathbb{R}} % Same thing as Ryan's \R
\newcommand{\Rd}{\Reals^d}
\newcommand{\wb}[1]{\widebar{#1}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\1}{\mathbf{1}}
\newcommand{\bj}{{\bf j}}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wc}{0}{mathx}{"71}
%%% End Alden's Additions

\begin{document}
	\begin{center} {\Large{\bf{Convergence Rates of Principal Components Regression with Laplacian Eigenmaps}}}

		\vspace*{.3cm}

		{\large{
				\begin{center}
					Alden Green~~~~~ Sivaraman Balakrishnan~~~~~ Ryan J. Tibshirani\\
					\vspace{.2cm}
				\end{center}

				\begin{tabular}{c}
					Department of Statistics and Data Science \\
					Carnegie Mellon University
				\end{tabular}

				\vspace*{.2in}

				\begin{tabular}{c}
					\texttt{\{ajgreen,siva,ryantibs\}@stat.cmu.edu}
				\end{tabular}
		}}

		\vspace*{.2in}

		\today
		\vspace*{.2in}
	\end{center}

	\section{Introduction}

	Laplacian Eigenmaps (LE)~\citep{belkin03a} is a method for nonlinear dimensionality reduction and data representation. Given scattered data points $\{X_1,\ldots,X_n\} \subset \Reals^d$, LE maps each $X_i$ to a vector $(v_{i,1},\ldots,v_{i,K})$ according to the following steps.
	\begin{enumerate}
		\item First, LE forms a \emph{neighborhood graph} $G = (V,W)$ over the points $\{X_1,\ldots,X_n\}$. The graph $G$ is an undirected, weighted graph, with vertices $V = \{X_1,\ldots,X_n\}$, and weighted edges $W_{ij}$ which correspond to the proximity between points $X_i$ and $X_j$, say in Euclidean distance.(For a formal definition see Section~\ref{})
		\item Next, LE forms an (unweighted) \emph{graph Laplacian} matrix $L \in \Reals^{n \times n}$, a symmetric and diagonally dominant matrix with diagonal elements $L_{ii} = \sum_{j = 1}^{n} W_{ij}$, and off-diagonal elements $L_{ij} = -W_{ij}$. 
		\item Finally, LE takes the eigendecomposition $L = \sum_{k = 1}^{n} \lambda_k v_k v_k^{\top}$, and outputs the vectors $(v_{1,i},\ldots,v_{K,i}) \in \Reals^K$ for each $i = 1,\ldots,n$.
	\end{enumerate}

	A natural way to use LE is by taking the vectors $\{(v_{i,1},\ldots,v_{i,K})\}_{i = 1}^{n}$ to be features in a downstream supervised learning algorithm. In this paper, we study a simple method along these lines: Principal Components Regression with Laplacian-Eigenmaps (PCR-LE), a method for nonparametric regression which operates by running ordinary least squares (OLS) using the features learned by LE. Given pairs of design points and responses $(X_1,Y_1),\ldots, (X_n,Y_n)$, PCR-LE computes an estimate $\wh{f} \in \Reals^n$,
	\begin{equation}
	\label{eqn:pcr-le}
	\wh{f} := \argmin_{f \in \mathrm{span}\{v_1,\ldots,v_K\}} \|{\bf Y} - f\|_2^2 = V_K V_K^{\top} {\bf Y},
	\end{equation}
	where ${\bf Y} = (Y_1,\ldots,Y_n) \in \Reals^n$ is the vector of responses, $\|\cdot\|_2$ denotes the usual Euclidean norm in $\Reals^n$, and the equality in \eqref{eqn:pcr-le} is due to the orthogonality of eigenvectors with respect to Euclidean inner product.

	LE has been practically very successful, and by now has been used for various statistical tasks such as spectral clustering, manifold learning, level-set estimation, semi-supervised learning, etc. At this point there exists a rich literature \citep{koltchinskii2000,belkin07,vonluxburg2008,burago2014,shi2015,singer2017,garciatrillos18,trillos2019, calder2019, cheng2021,dunson2021} explaining this practical success from a theoretical perspective. Loosely speaking, these works model the design points as being independent samples from a distribution $P$ with density $p$, and show that in this case the eigenvectors of the graph Laplacian $L$ are good empirical approximations of population-level objects. These population-level objects are the eigenfunctions $\psi_k$---meaning solutions, along with eigenvalues $\rho_k$, to the equation $\Delta_P \psi_k = \rho_k \psi_k$--- of the density-weighted Laplacian operator
	\begin{equation}
	\label{eqn:density-weighted-laplace}
	\Delta_Pf := -\frac{1}{p}~ \mathrm{div}(p^2 \nabla f).
	\end{equation}  
	(Here $\mathrm{div}$ stands for the divergence operator, and $\nabla$ for the gradient.) These eigenfunctions in turn characterize various interesting structural aspects of $p$, such as the location and number of high- and low-density regions, the shape and intrinsic dimension of its support, and so forth.

	These aforementioned works explain what kind of representation LE learns, and the accuracy with which it learns this representation. However, this theory does not address the convergence rates of PCR-LE. That is the major question we answer in this paper. We adopt the classical model of nonparametric regression with random design: we observe independent pairs $(X_1,Y_1),\ldots,(X_n,Y_n)$ of design points and responses, where the design points $\{X_1,\ldots,X_n\}$ are sampled independently from an unknown distribution $P$ supported on $\mc{X} \subseteq \Rd$, and the responses follow a signal plus Gaussian noise model,
	\begin{equation}
	\label{eqn:model}
	Y_i = f_0(X_i) + w_i, \quad w_i \sim N(0,1),
	\end{equation}
	with noise $w_i$ independent of $X_i$. The task is to learn the regression function $f_0$, which is unknown but assumed to belong to a Sobolev space $H^s(\mc{X})$. We consider two settings: one where $\mc{X}$ is a full-dimensional domain, and the other where $\mc{X}$ is a low-dimensional submanifold of $\Rd$. In each setting, we derive upper bounds which imply that the PCR-LE estimate $\wh{f}$, and a test using the statistic $T = \|\wh{f}\|_2^2$, are statistically optimal methods for two classical problems in nonparametric regression: estimation and goodness-of-fit testing.

	At first glance, the optimality of PCR-LE is somewhat surprising. As a function of the sample size $n$, known rates of convergence of LE are much slower than the minimax-optimal rates of convergence over $H^s(\mc{X})$ (roughly speaking $n^{-2s/(2s + d)}$ for estimation and $n^{-4s/(4s + d)}$ for testing). This is not due to suboptimal theory for LE, but rather reflects a fundamental fact: regression using learned features does not rely on accurately learning the features. It is essential to employ modes of analysis which exploit this fact in order to derive sharp rates of convergence for PCR-LE. We explain this in more detail after summarizing our main results.

	\paragraph{Sobolev spaces and spectral series regression.}
	To analyze PCR-LE, we work in a classical situation where the regression function is assumed to belong to a (Hilbert-)Sobolev space. For an open domain $\mc{X} \subseteq \Rd$, the Sobolev space $H^s(\mc{X})$ consists of all functions $f \in L^2(\mc{X})$ which are $s$-times weakly differentiable, with all order-$s$ partial derivatives $D^{\alpha}f \in L^2(\mc{X})$. We study regression over Sobolev spaces in part because generally speaking, the minimax rates are well-understood; as mentioned before, as a function of the sample size $n$ they are $n^{-2s/(2s + d)}$ for estimation, and $n^{-4s/(4s + d)}$ for testing. For this reason, regression over Sobolev spaces is a good setting in which to see whether PCR-LE measures up to simpler and more classical minimax-optimal approaches, which have strong theoretical guarantees but are less often used in practice.

	Moreover, we view PCR-LE as being particularly well-suited for regression over Sobolev spaces, due to their close connection with \emph{spectral series regression}. Spectral series regression computes empirical Fourier coefficients and truncates to the lowest frequency eigenfunctions, producing the estimate
	\begin{equation}
	\label{eqn:population-level_spectral_series}
	\wt{a}_k = \frac{1}{n}\sum_{i = 1}^{n} Y_i \psi_k(X_i), \quad \wt{f}(x) = \sum_{k = 1}^{K} \wt{a}_k \psi_k(x).
	\end{equation} 
	Under appropriate boundary conditions, the Sobolev spaces $H^s(\mc{X})$ can roughly be characterized as consisting of those functions $f = \sum_{k} a_k \psi_k \in L^2(\mc{X})$ for which the generalized Fourier coefficients $\{a_k\}_{k = 1}^{\infty}$ satisfy the decay condition $\sum_{k} a_k \rho_k^2 < \infty$. Heuristically, this decay condition justifies the truncation in~\eqref{eqn:population-level_spectral_series}---since the truncation incurs only a limited amount of bias for any $f_0 \in H^s(\mc{X})$---and for this reason spectral series regression over Sobolev spaces has been studied since \textcolor{red}{(?)}, shown to be optimal for estimation by \textcolor{red}{(?)}, and for goodness-of-fit testing by \textcolor{red}{(?)}.

	PCR-LE serves as an empirical approximation to spectral series regression, since the eigenvectors $v_k$ are empirical approximations to eigenfunctions $\psi_k$. Viewed in this light, a major advantage of PCR-LE is that it operates independently of the design distribution $P$. In the contrast, the spectral series estimator defined in~\eqref{eqn:population-level_spectral_series} relies on diagonalizing the density-weighted Laplacian $\Delta_P$. In our context, where $P$ is treated as unknown,~\eqref{eqn:population-level_spectral_series} must be viewed as an oracle method, and to emphasize this  we henceforth refer to~\eqref{eqn:population-level_spectral_series} as \emph{population-level spectral series regression}. On the other hand, naturally PCR-LE incurs some extra error by using an empirical approximation to the underlying basis $\{\psi_k\}$, and our work shows that in many cases, this extra error is small enough that it does not change the overall rate of convergence.

	\subsection{Main contributions}
	Summarized succinctly, our main contribution is to theoretically analyze nonparametric regression with PCR-LE and establish upper bounds which imply that, in many cases, this method achieves optimal rates of convergence over Sobolev spaces.

	\paragraph{Rates of convergence: population-level spectral series regression.}
	As we have already mentioned, the minimax-optimal rates over Sobolev spaces are generally well-known, as are upper bounds for population-level spectral series methods which match these rates. However, we could not find precisely stated results applying to our setting, where in particular
	\begin{enumerate}
		\item We consider subcritical Sobolev spaces $H^s(\mc{X})$, where the smoothness parameter $s$ satisfies $s < d/2$ and so $H^s(\mc{X})$ does not continuously embed into the space of continuous functions $C^0(\mc{X})$, and 
		\item We consider general design distributions $P$, which may satisfy certain regularity conditions but are not limited to being, say, the uniform distribution over $[0,1]^d$. 
	\end{enumerate}
	For completeness, we analyze population-level spectral series methods in this general setting, and establish upper bounds showing that such methods converge at the ``usual'' rates of $n^{-2s/(2s + d)}$ for estimation and $n^{-4s/(4s + d)}$ for testing. This analysis relies heavily on certain asymptotic properties of the continuum eigenfunctions $\psi_k$ and eigenvalues $\rho_k$, which hold for quite general second-order differential operators $\mc{L}$ including the density-weighted Laplacian $\mc{L} = \Delta_P$.

	\paragraph{Rates of convergence: PCR-LE.}
	The rest of our results consist of various upper bounds on the rates of convergence for the PCR-LE estimator $\wh{f}$, and a test using the statistic $T = \|\wh{f}\|_2^2$. These upper bounds quantify two important properties of PCR-LE: first, that it can take advantage of smooth higher-order derivatives, and second that it can adapt to low intrinsic dimension of the design distribution, each in an optimal manner. We first model the design distribution $P$ as having support $\mc{X}$ which is a full-dimensional set in $\Rd$. In this case, our main contributions are as follows:
	\begin{itemize}
		\item  Over a ball in the Sobolev space $H^{s}(\mc{X})$, we establish that the PCR-LE estimator $\wh{f}$ has in-sample mean-squared error of at most on the order of $n^{-2s/(2s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d$. 
		\item We show that a test based on the statistic $\|\wh{f}\|_2^2$ has a squared critical radius on the order of $n^{-4s/(4s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d \in \{1,2,3,4\}$. 
	\end{itemize}
	We then consider the behavior of PCR-LE when the data satisfies a \emph{manifold hypothesis}, meaning the design distribution is supported on an (unknown) domain $\mc{X}$ which is a submanifold of $\Rd$ of intrinsic dimension $m \in \mathbb{N}, m < d$. In this case, our upper bounds imply that:
	\begin{itemize}
		\item When $f_0 \in H^s(\mc{X})$ the PCR-LE estimator has in-sample mean squared error of at most $n^{-2s/(2s + m)}$, when $s \in \{1,2,3\}$ and for any $m \in \mathbb{N}$. 
		\item A test based on the statistic $\|\wh{f}\|_2^2$ has a squared critical radius on the order of $n^{-4s/(4s + m)}$, when $s \in \{1,2,3\}$ and $m \in \{1,2,3,4\}$.
	\end{itemize}
	We note that under the manifold hypothesis, to the best of our knowledge the minimax rates for Sobolev spaces $H^s(\mc{X})$ under the assumption of random design with design distribution supported on an \emph{unknown} manifold have not been worked out. Our upper bounds confirm that these rates are the same as for regression over a known manifold, at least for the values of $s$ and $m$ mentioned above.

	In all these cases, our bounds also depend optimally on the radius $M$ of the Sobolev ball under consideration. However, for some values of $s$ (number of derivatives) and $d$ (dimension), there do exist gaps between our upper bounds on the error of PCR-LE and the minimax rates. Although we do not give corresponding lower bounds verifying the tightness of our analysis, we believe these gaps reflect the true behavior of the method rather than some looseness in our analysis, and we comment more on this at relevant parts in the text. For completeness, we summarize all of our upper bounds---those which match the minimax rates, and those which do not---in Tables~\ref{tbl:estimation_rates} and~\ref{tbl:testing_rates}.
	\begin{table}
		\begin{center}
			\begin{tabular}{p{.2\textwidth} | p{.14\textwidth} p{.12\textwidth} }
				Smoothness order & Flat Euclidean (Model~\ref{def:model_flat_euclidean}) & Manifold (Model~\ref{def:model_manifold}) \\
				\hline
				$s \leq 3$ & ${\bf n^{-2s/(2s + d)}}$ & ${\bf n^{-2s/(2s + m)}}$ \\
				$s > 3$  & ${\bf n^{-2s/(2s + d)}}$ & $n^{-6/(6 + m)}$
			\end{tabular}
		\end{center}
		\caption{Summary of Laplacian eigenmaps estimation rates over Sobolev balls. Bold font marks minimax optimal rates. In each case, rates hold for all $d \in \mathbb{N}$ (under Model~\ref{def:model_flat_euclidean}), and for all $m \in \mathbb{N}, 1 < m < d$ (under Model~\ref{def:model_manifold}). Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps estimator is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal.}
		\label{tbl:estimation_rates}
	\end{table}

	\begin{table}
		\begin{center}
			\begin{tabular}{p{.175\textwidth} p{.175\textwidth} | p{.14\textwidth} p{.12\textwidth} }
				Smoothness order & Dimension & Flat Euclidean (Model~\ref{def:model_flat_euclidean}) & Manifold (Model~\ref{def:model_manifold}) \\
				\hline
				\multirow{2}{*}{$s = 1$} & $\dim(\mc{X}) < 4$ & ${\bf n^{-4s/(4s + d)}}$ & ${\bf n^{-4s/(4s + m)}}$ \\
				& $\dim(\mc{X}) \geq 4$ & ${\bf n^{-1/2}}$ & ${\bf n^{-1/2}}$ \\
				\hline
				\multirow{3}{*}{$s = 2$ or $3$} & $\dim(\mc{X}) \leq 4$  & ${\bf n^{-4s/(4s + d)}}$ & ${\bf n^{-4s/(4s + m)}}$ \\
				& $4 <\dim(\mc{X}) < 4s$  & $n^{-2s/(2(s - 1) + d)}$ & $n^{-2s/(2(s - 1) + m)}$\\
				& $\dim(\mc{X}) \geq 4s$ & ${\bf n^{-1/2}}$ & ${\bf n^{-1/2}}$ \\
				\hline
				\multirow{3}{*}{$s > 3$} & $\dim(\mc{X}) \leq 4$ & ${\bf n^{-4s/(4s + d)}}$ & $n^{-12/(12 + d)}$ \\
				& $4 < \dim(\mc{X}) < 4s$ & $n^{-2s/(2(s - 1) + d)}$ & $n^{-6/(4 + m)}$ \\
				& $\dim(\mc{X}) \geq 4s$ & ${\bf n^{-1/2}}$ & ${\bf n^{-1/2}}$ \\
			\end{tabular}
		\end{center}
		\caption{Summary of Laplacian eigenmaps testing rates over Sobolev balls. Bold font marks minimax optimal rates. Rates when $d > 4s$ assume that $f_0 \in L^4(\mc{X})$, and depend on $\|f_0\|_{L^4(\mc{X})}$. Although we suppress it for simplicity, in all cases when othe Laplacian eigenmaps test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal.}
		\label{tbl:testing_rates}
	\end{table}

	\paragraph{Technical contributions.}
	One of our primary technical contributions is a strategy for analyzing regression methods using learned features which leads to sharp rates of convergence.

	At an abstract level PCR-LE is a two-stage algorithm: the first stage (LE) uses the observed design points $\{X_1,\ldots,X_n\}$ to learn a feature representation ($X_i \mapsto (v_{1,i},\ldots,v_{K,i})$) which is an empirical approximation to some ideal population-level representation ($X_i \mapsto (\psi_{1}(X_i),\ldots,\psi_{K}(X_i))$);  the second stage then applies a simple method (OLS) to the learned features, and obtains an estimate. At first blush, one might expect the error for PCR-LE to be likewise decomposed into two parts: first, the error with which the features are learned (i.e. error due to sampling variation in $X$), and the second being error with which, given ideal population-level features, the function is learned (i.e. error due to noise in $Y$.)

	Crucially, our analysis \emph{does not} work in this way. This is important because, as already mentioned, all known upper bounds on the error with which the learned features (Laplacian eigenvectors) approximate their population-level limits (continuum Laplacian eigenfunctions) are much larger than the minimax rates of convergence over $H^s(\mc{X})$. For example,
	\begin{itemize}
		\item The best known rates of convergence for graph Laplacian eigenvectors (due to~\cite{cheng2021}) are
		\begin{equation}
		\label{eqn:eigenvector_convergence}
		\max_{1 \leq k \leq K}\|\sqrt{n}v_k - \psi_k\|_n^2 \leq C_{K} n^{-2/(4 + d)},
		\end{equation}
		where the constant $C_K$ depends on $K$ in an unknown manner.
		\item The dependence of $C_K$ on $K$ is relevant for our interests where $K$ is growing with $n$. The earlier works of \cite{burago2014,trillos2019} give upper bounds which, although implying weaker rates of convergence as a function of $n$, do keep track of $C_K$, and suggest that it should depend on the inverse of the spectral gap 
		\begin{equation}
		\label{eqn:spectral_gap}
		C_K = \frac{C}{\rho_{K + 1} - \rho_K} \geq C K^{1 - 2/d},
		\end{equation}
		where $C$ is a constant depending only on the dimension $d$, and the latter inequality follows from the Weyl's Law scaling $\rho_K \asymp K^{2/d}$.
	\end{itemize}
	Although these upper bounds may not reflect the true rate of convergence of graph Laplacian eigenvectors---this is still an active area of research, and no lower bounds are known---it seems very unlikely that the true rate matches even the estimation minimax rate $n^{-2s/(2s + d)}$, which after all approaches the dimension-free rate $1/n$ for large values of $s$.

	Instead of relying on convergence of eigenvectors to eigenfunctions, our analysis proceeds via a bias-variance decomposition at the level of the graph. Focusing for simplicity on estimation, this is 
	\begin{equation}
	\label{eqn:bias_variance_estimation}
	\|f_0 - \wh{f}\|_n^2 \leq 2\bigl(\|f_0 - V_KV_K^{\top}f_0\|_n^2 + \|V_KV_K^{\top}({\bf Y} - f_0)\|_n^2\bigr) 
	\end{equation}
	The second term in~\eqref{eqn:bias_variance_estimation} is the variance, and as usual for OLS estimates depends only on the degrees of freedom $\mathrm{df}(\wh{f}) = \mathrm{tr}(V_KV_K^{\top}) = K$. More surprisingly, the first term (squared-bias) can also be upper bounded without appealing to the eigenfunctions $\psi_k$. Letting $S = L^{\dagger}$ be the pseudo-inverse of the graph Laplacian,
	\begin{equation}
	\begin{aligned}
	\|(I - V_KV_K^{\top})f_0\|_n^2 & = \|(I - V_KV_K^{\top})S^{s/2}L^{s/2}f_0\|_n^2 \leq \|(I - V_KV_K^{\top})S^{s/2}\|_{op}^2 \|L^{s/2}f_0\|_n^2 = \frac{f_0^{\top} L^s f_0}{n \lambda_{K + 1}^{s}}.
	\end{aligned}
	\end{equation}
	Thus we obtain an upper bound on the in-sample sample mean squared error $\|\wh{f} - f_0\|_n^2$ that is \emph{independent of the error with which graph Laplacian eigenvectors $v_k$ approximate population-level eigenfunctions $\psi_k$}.

	Instead, our upper bound on the error of PCR-LE is determined by a pair of graph functionals: the quadratic form $f_0^{\top}L^s f_0$, and the graph Laplacian eigenvalue $\lambda_{K + 1}$. Theoretically speaking, this brings two advantages over directly analyzing convergence of eigenvectors to eigenfunctions. First, the graph functionals can be more tightly controlled than the eigenvectors $v_k$; for example, the error  Moreover, in order to obtain rates of convergence we do not require that these functionals themselves converge to population-level limits, but only that they be stochastically bounded on the right order. To derive our ultimate upper bounds we use some existing results regarding neighborhood graph Laplacians, and prove some new ones which may be of independent interest.

	A last observation regarding the difference between feature learning and regression using learned features. To balance the bias and variance terms in~\eqref{eqn:bias_variance_estimation}, we will end up using $K(n) := n^{d/(2s + d)}$ eigenvectors. Although this is the usual choice for spectral series regression over order-$s$ Sobolev spaces, in the context of PCR-LE it leads to a striking conclusion. Namely, examining~\eqref{eqn:eigenvector_convergence} and~\eqref{eqn:spectral_gap}, we see that for all dimensions $d \geq 3$, known upper bounds on $\|\sqrt{n} v_{K(n)} - \psi_{K(n)}\|_n^2$ do not converge to $0$ as $n \to \infty$. In words, it is not merely the case that PCR-LE converges at a faster rate than LE itself, but that PCR-LE can profitably use eigenvectors for regression which may be inconsistent estimates of their population-level limits.

	To summarize, our work demonstrates, broadly speaking, that regression using learned features can be analyzed independent of the accuracy of the features themselves. Regression using learned features is a general and widely applied paradigm, and we believe this observation may have consequences outside of its application to PCR-LE in this work.

	\subsection{Related work}

	Much of the work regarding regression using neighborhood graph Laplacians deals with \emph{semi-supervised learning}, where in addition to the labeled data $(X_1,Y_1),\ldots,(X_n,Y_n)$ one observes unlabeled points $(X_{n + 1},\ldots,X_{N})$. The landmark paper of \cite{zhu2003semisupervised} proposed to interpolate the observed values by~\emph{harmonic extension}, i.e. compute $L_N$ the Laplacian matrix corresponding to a graph formed over all design points $X_1,\ldots,X_N$, and then solve the constrained problem
	\begin{equation*}
	\minimize_{f \in \Reals^N} f^{\top} L_N f \quad \mathrm{subject\,\,to}~~~ f_i = Y_i~~\textrm{for $i = 1,\ldots,n$.}
	\end{equation*}
	Conventional wisdom says that harmonic extension is sensible only when the responses are noiseless, $Y_i = f_0(X_i)$, and that in the noisy setting one should instead solve the penalized formulation
	\begin{equation}
	\label{eqn:graph_laplacian_regularization_ssl}
	\wc{f} = \argmin_{f \in \Reals^N} \|{\bf Y} - f\|_n^2 + \lambda f^{\top} L_N f.
	\end{equation}
	Notwithstanding their intuitive appeal, both the constrained and penalized problems have issues when $d > 1$ and $n/N \to 0$~\citep{nadler09,calder2019b, calder2020}, and in certain cases the estimates are degenerate, meaning they are ``spiky'' at labeled data points and close to constant everywhere else. One solution to this problem is to instead use Laplacian Eigenmaps for semi-supervised learning (SSL-LE), i.e. compute the eigendecomposition $L_N = \sum_{k = 1}^{N} \lambda_k u_k u_k^{\top}$ and solve the problem
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_ssl}
	\wb{f} := \argmin_{f \in \mathrm{span}\{u_1,\ldots,u_K\}} \|{\bf Y} - f\|_n^2.
	\end{equation}
	\cite{zhou2011,lee2016} analyze SSL-LE in a particular asymptotic regime where $n$ is held fixed while $N \to \infty$, and show that it achieves minimax-optimal rates of convergence over $H^s(\mc{X})$ as a function of $n$, the number of labeled points. Of course, in this regime the relevant eigenvectors of $L_n$ all converge to their continuum limits, meaning
	\begin{equation*}
	\lim_{N \to \infty} \max_{1 \leq K \leq K(n)} \|\sqrt{N} u_k - \psi_k\|_N^2 = 0,
	\end{equation*}
	and consequently $\lim_{N \to \infty} \|\wb{f} - \wt{f}\|_N^2 = 0$. Thus the analysis of SSL-LE  reduces to that of population-level spectral series regression. The supervised setting (where $N = n$) we consider in this work is very different, and analyzing PCR-LE necessitates an entirely different approach, as explained in the preceding section.

	There has been much less work, relatively speaking, regarding regression with neighborhood graph Laplacians in the supervised setting, which is the focus of this work. \citet{lee2016} also analyze a variant of PCR-LE, but derive suboptimal rates of convergence. \citet{trillos2020} study nonparametric regression via graph Laplacian regularization, where the estimator is computed by solving~\eqref{eqn:graph_laplacian_regularization_ssl}, but with no unlabeled data. They establish the upper bound $\max_{i = 1,\ldots,n}|\wc{f}(X_i) - f_0(X_i)| \leq C n^{-2/(2 + d)}$ under the assumption $f_0 \in C^2(\mc{X})$, which is slower than the minimax rate $n^{-4/(4 + d)}$ for this class. In a previous paper~\citep{green2021}, we (the authors) also considered nonparametric regression via graph Laplacian regularization, when the regression function $f_0 \in H^1(\mc{X})$ and loss is measured using in-sample mean squared error. We found that graph Laplacian regularization is consistent as $n \to \infty$, and optimal (using mean-squared error) over the first-order Sobolev space $H^1(\mc{X})$ when $d \in \{1,2,3,4\}$. However, graph Laplacian regularization neither takes advantage of smooth higher-order derivatives, nor is it provably optimal over $H^1(\mc{X})$ for all dimensions $d$. One of our motivations for considering PCR-LE was to find an estimator which addressed these deficiencies. In this work we indeed find that PCR-LE has much stronger optimality properties than graph Laplacian regularization.

	Most work on supervised learning using graphs adopts a \emph{fixed design} perspective, treating the design points $X_1 = x_1,\ldots,X_n = x_n$ as vertices of a fixed graph, and carrying out inference with respect to the conditional mean vector $(f_0(x_1),\ldots,f_0(x_n))$. In this setting, matching upper and lower bounds have been established that certify the optimality of graph-based methods for estimation \citep{wang2016,hutter2016,sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}) and testing \citep{sharpnack2010identifying,sharpnack2013b,sharpnack2013,sharpnack2015} over different ``function'' classes (in quotes because these classes really model the $n$-dimensional vector of evaluations). This setting is quite general, because the graph need not be a geometric graph defined on a vertex set which belongs to Euclidean space. On the other hand, depending on the data collection process, it may be unnatural to model the design points as being a priori fixed, and the estimand as being a vector which exhibits a discrete notion of ``smoothness'' over this fixed design. Instead, we adopt the \emph{random design} perspective, and seek to estimate a function that we assume exhibits a more classical notion of smoothness.

% Bibliography
\bibliographystyle{plainnat}
\bibliography{../../../graph_regression_bibliography}

\end{document}

### 2021-10-16 manuscripts/laplacian_eigenmaps/arxiv/laplacian_eigenmaps_regression.tex
Updated Laplacian Eigenmaps manuscript.

@@
 				\vspace{.2cm}
-				via Laplacian Eigenmaps with Neighborhood Graphs}}}
+				via Laplacian Eigenmaps with Neighborhood Graphs} \\}
+			\textcolor{red}{(TODO): Keep this title?}}
 
@@
 \begin{abstract}
-	In this paper we study the statistical properties of nonparametric regression methods based on \emph{Laplacian eigenmaps}.
-These are spectral projection methods, which involve projecting a vector of observed responses ${\bf Y} = (Y_1,\ldots,Y_n)$ onto a subspace spanned by certain eigenvectors of a neighborhood graph Laplacian.
-We show that in many cases these methods achieve minimax rates of convergence over continuum function classes.
-Specifically, we consider the problem of random design regression over Sobolev spaces $H^s(\mc{X})$---where $\mc{X} \subseteq \Rd$ is the support of the design density $p$---and show that under sufficient smoothness conditions on $p$, methods based on Laplacian eigenmaps achieve the optimal rates for both estimation (where the optimal rate is known to be $n^{-2s/(2s + d)}$) and goodness-of-fit testing ($n^{-4s/(4s + d)}$).
-We also show that Laplacian eigenmaps is \emph{manifold adaptive}: that is, we consider the situation where $\mc{X}$ is a manifold of small intrinsic dimension $m$, and give upper bounds establishing that Laplacian eigenmaps achieves the faster minimax estimation ($n^{-2s/(2s + m)}$) and testing ($n^{-4s/(4s + m)}$) rates of convergence.
-Finally, since the Laplacian eigenmaps estimator is defined only at observed design points, we propose an out-of-sample extension based on kernel smoothing, and prove that it achieves optimal out-of-sample error.
+	In this paper we study the statistical properties of Principal Components Regression with Laplacian Eigenmaps (PCR-LE), a method for nonparametric regression based on Laplacian Eigenmaps (LE).
+PCR-LE works by projecting a vector of observed responses ${\bf Y} = (Y_1,\ldots,Y_n)$ onto a subspace spanned by certain eigenvectors of a neighborhood graph Laplacian.
+We show that in many cases PCR-LE achieves minimax rates of convergence over continuum function classes.
+Specifically, we consider the problem of random design regression over Sobolev spaces $H^s(\mc{X})$---where $\mc{X} \subseteq \Rd$ is the support of the design distribution $P$---and show that under sufficient smoothness conditions on the design density $p$, PCR-LE achieves the optimal rates for both estimation (where the optimal rate is known to be $n^{-2s/(2s + d)}$) and goodness-of-fit testing ($n^{-4s/(4s + d)}$).
+We also show that PCR-LE is \emph{manifold adaptive}: that is, we consider the situation where $\mc{X}$ is a manifold of small intrinsic dimension $m$, and give upper bounds establishing that PCR-LE achieves the faster minimax estimation ($n^{-2s/(2s + m)}$) and testing ($n^{-4s/(4s + m)}$) rates of convergence.
+Interestingly, these rates are almost always much faster than the known rates of convergence of graph Laplacian eigenvectors to their population-level limits; in other words, for this problem regression with learned features appears to be much easier, statistically speaking, than feature learning.
 We support these theoretical results with empirical evidence.

### 2021-10-16 manuscripts/laplacian_eigenmaps/main_text.tex
Updated Laplacian Eigenmaps manuscript.

@@
 \label{sec:introduction}
-Two fundamental statistical goals are to learn and leverage structure of an unknown distribution on the basis of samples.
-Neighborhood graph Laplacians, and their associated spectra, have proven to be powerful and general tools for both purposes.
-They have been successfully used for various statistical tasks such as (spectral) clustering, manifold learning, level-set estimation, semi-supervised learning, etc.
-
-By now there exists a rich literature \citep{koltchinskii2000,belkin07,vonluxburg2008,burago2014,shi2015,singer2017,garciatrillos18,trillos2019, calder2019, cheng2021,dunson2021} explaining this practical success from a theoretical perspective.
-Loosely speaking, these works justify graph spectral methods by showing that the eigenvectors and eigenvalues of graph Laplacians are empirical approximations of population-level objects: the eigenvalues and eigenfunctions of a density-weighted Laplacian operator.
-These population-level objects in turn characterize various structural aspects of the density, such as its high- and low-density regions, the shape and intrinsic dimension of its support, and so forth.
-Together, these insights give a reasonably complete explanation of how graph spectra \emph{learn} the structure of an underlying distribution.
-In contrast, in this paper we develop theory showing that graph spectral methods can effectively \emph{leverage} learned structure to perform downstream (supervised) learning tasks.
-Our main goal is to show that spectral projection procedures---which use eigenvectors of a neighborhood graph Laplacian as features in linear regression---are statistically optimal methods for two problems in nonparametric regression: estimation and goodness-of-fit testing.
-
-The regression methods we consider are based on \emph{Laplacian eigenmaps}, first introduced by~\cite{belkin03a}.\footnote{Technically speaking, as used by~\cite{belkin03a} ``Laplacian eigenmaps'' refers to a spectral embedding procedure.
-What we call Laplacian eigenmaps is a two-step regression method: the first step is this spectral embedding, and the second step computes the estimator $\wh{f}$ by running ordinary least squares in embedding space.
-Since in this work we only ever consider the embedding within the context of this two-step regression method, we take the liberty of referring to the entire procedure as Laplacian eigenmaps.} Given samples $(X_1,Y_1),\ldots,(X_n,Y_n)$, Laplacian eigenmaps first forms a neighborhood graph, with vertices at the design points $\{X_1,\ldots,X_n\}$ and weighted edges between pairs of sufficiently close points $X_i$ and $X_j$.
-The Laplacian eigenmaps estimator $\wh{f}$ is then computed using the spectral decomposition of an (unweighted) Laplacian defined over this neighborhood graph, by projecting the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto a subspace spanned by a certain number of ``low-frequency'' graph Laplacian eigenvectors; more precisely, the eigenvectors corresponding to the smallest eigenvalues of the graph Laplacian.
-The Laplacian eigenmaps test statistic $\wh{T} = \|\wh{f}\|_n^2$ can in turn be used as a statistic to detect whether any signal is present, that is, whether the regression function $f_0(x) = \mathbb{E}[Y|X = x]$ is anywhere non-zero.
-(The precise definition of these methods is given in Section~\ref{subsec:laplacian_eigenmaps}.) We study these methods through the lens of a minimax analysis.
-Our major contribution is to provide upper bounds on the error of the Laplacian eigenmaps estimator and test which imply that they achieve the minimax optimal rates of convergence over natural smoothness classes.
-
-More specifically, we consider the worst-case risk of Laplacian eigenmaps over normed balls in Sobolev spaces.
-Roughly speaking, for a domain $\mc{X} \subseteq \Rd$, the order-$s$ (Hilbert-)Sobolev space $H^s(\mc{X})$ consists of those functions which are $s$-times weakly differentiable, with order-$1$ up through order-$s$ derivatives bounded in $L^2(\mc{X})$ norm.
-Equivalently, Sobolev spaces can be defined in a spectral manner;  letting $\psi_1,\psi_2,\ldots$ be the eigenfunctions of a Laplacian operator defined with respect to $\mc{X}$, Sobolev spaces consist of those functions $f$ whose Fourier coefficients $\theta_k := \int_{\mc{X}} f \psi_k$ are bounded in a particular sequence norm (see Section~\ref{subsec:spectral_projection}).
-The minimax estimation and testing rates for Sobolev spaces are generally speaking well-known---when using mean-squared error as the loss, the estimation rate is $n^{-2s/(2s + d)}$, and the testing rate is $n^{-4s/(4s + d)}$---and are achieved by various methods.
-However, since Sobolev spaces admit a spectral characterization in terms of $\psi_1,\psi_2,\ldots$, spectral projection methods that use these eigenfunctions as features are a particularly natural choice for regression over Sobolev spaces, and are in a sense canonical estimators for these function classes.
-% (AG 8/13/21): Last clause makes a strong statement that I personally believe, but may irritate certain readers. Keep or remove?
-% (AG 8/18/21): I do not love the name ``classical spectral projection method''. It doesn't roll off the tongue, and ``classical'' doesn't really convey what the method is. Is there a better name?
-
-Laplacian eigenmaps serves as a data-dependent approximation to these population-level spectral projection methods.\footnote{We use the term ``population-level'' to emphasize that these methods use eigenfunctions which are the infinite-data limits of graph Laplacian eigenvectors.} This data dependency can be very useful.
-As the name suggests, population-level spectral methods require a priori knowledge of the design distribution, since they are defined with respect to the eigenfunctions of a Laplacian operator which itself depends on this distribution.
-For this reason, these methods are typically studied only under restrictive conditions on the design; for instance, that the design points are uniformly distributed over the unit cube.
-In many situations such restrictive conditions are not satisfied, and it may not be reasonable to assume prior knowledge of the design distribution.
-In contrast to population-level methods, Laplacian eigenmaps implicitly learns structure from the design points $\{X_1,\ldots,X_n\}$, and can thus be used even when the design distribution is unknown.
-Of course, this comes with a trade off.
-Intuitively, using an empirical approximation to the ``right'' population-level basis will incur some additional error.
-The question is whether, notwithstanding this additional error, Laplacian eigenmaps shares the strong statistical properties of its population-level counterpart.
-
-We find that in many cases, the answer is yes: under appropriate regularity conditions, when correctly tuned Laplacian eigenmaps is minimax optimal over Sobolev spaces.
-Specifically, our main contributions are as follows:
+
+Laplacian Eigenmaps (LE)~\citep{belkin03a} is a method for nonlinear dimensionality reduction and data representation.
+Given scattered data points $\{X_1,\ldots,X_n\} \subset \Reals^d$, LE maps each $X_i$ to a vector $(v_{i,1},\ldots,v_{i,K})$ according to the following steps.
+\begin{enumerate}
+	\item First, LE forms a \emph{neighborhood graph} $G = (V,W)$ over the points $\{X_1,\ldots,X_n\}$.
+The graph $G$ is an undirected, weighted graph, with vertices $V = \{X_1,\ldots,X_n\}$, and weighted edges $W_{ij}$ which correspond to the proximity between points $X_i$ and $X_j$, say in Euclidean distance.
+	\item Next, LE forms an (unweighted) \emph{graph Laplacian} matrix $L \in \Reals^{n \times n}$, a symmetric and diagonally dominant matrix with diagonal elements $L_{ii} = \sum_{j = 1}^{n} W_{ij}$, and off-diagonal elements $L_{ij} = -W_{ij}$.
+	\item Finally, LE takes the eigendecomposition $L = \sum_{k = 1}^{n} \lambda_k v_k v_k^{\top}$, and outputs the vectors $(v_{1,i},\ldots,v_{K,i}) \in \Reals^K$ for each $i = 1,\ldots,n$.
+\end{enumerate} 
+A natural way to use LE is by taking the vectors $\{(v_{i,1},\ldots,v_{i,K})\}_{i = 1}^{n}$ to be features in a downstream supervised learning algorithm.
+In this paper, we study a simple method along these lines: Principal Components Regression with Laplacian-Eigenmaps (PCR-LE), a method for nonparametric regression which operates by running ordinary least squares (OLS) using the features learned by LE.
+Given pairs of design points and responses $(X_1,Y_1),\ldots, (X_n,Y_n)$, PCR-LE computes an estimate $\wh{f} \in \Reals^n$,
+\begin{equation}
+\label{eqn:pcr-le}
+\wh{f} := \argmin_{f \in \mathrm{span}\{v_1,\ldots,v_K\}} \|{\bf Y} - f\|_2^2
+\end{equation}
+where ${\bf Y} = (Y_1,\ldots,Y_n) \in \Reals^n$ is the vector of responses and $\|\cdot\|_2$ denotes the usual Euclidean norm in $\Reals^n$.
+(For a formal definition of LE and PCR-LE, see Section~\ref{subsec:laplacian_eigenmaps}.)
+
+LE has been practically very successful, and by now has been used for various statistical tasks such as spectral clustering, manifold learning, level-set estimation, semi-supervised learning, etc.
+At this point there exists a rich literature \citep{koltchinskii2000,belkin07,vonluxburg2008,burago2014,shi2015,singer2017,garciatrillos18,trillos2019, calder2019, cheng2021,dunson2021} explaining this practical success from a theoretical perspective.
+Loosely speaking, these works model the design points as being independent samples from a distribution $P$ with density $p$, and show that in this case the eigenvectors of the graph Laplacian $L$ are good empirical approximations of population-level objects.
+These population-level objects are the eigenfunctions $\psi_k$---meaning solutions, along with eigenvalues $\rho_k$, to the equation $\Delta_P \psi_k = \rho_k \psi_k$--- of the density-weighted Laplacian operator
+\begin{equation}
+\label{eqn:density-weighted-laplace}
+\Delta_Pf := -\frac{1}{p}~ \mathrm{div}(p^2 \nabla f).
+\end{equation}  
+(Here $\mathrm{div}$ stands for the divergence operator, and $\nabla$ for the gradient.) These eigenfunctions in turn characterize various interesting structural aspects of $p$, such as the location and number of high- and low-density regions, the shape and intrinsic dimension of its support, and so forth.
+
+These aforementioned works explain what kind of representation LE learns, and the accuracy with which it learns this representation.
+However, this theory does not address the convergence rates of PCR-LE.
+That is the major question we answer in this paper.
+We adopt the classical model of nonparametric regression with random design, where one observes independent pairs $(X_1,Y_1),\ldots,(X_n,Y_n)$ of design points and responses.
+We assume the design points $\{X_1,\ldots,X_n\}$ are sampled from an unknown distribution $P$ supported on $\mc{X} \subseteq \Rd$, and the responses follow a signal plus Gaussian noise model,
+\begin{equation}
+\label{eqn:model}
+Y_i = f_0(X_i) + w_i, \quad w_i \sim N(0,1),
+\end{equation}
+with noise variables $w_i$ independent of $X_i$.
+The task is to learn the regression function $f_0$, which is unknown but assumed to belong to a Sobolev space $H^s(\mc{X})$.
+We consider two settings: one where $\mc{X}$ is a full-dimensional domain, and the other where $\mc{X}$ is a low-dimensional submanifold of $\Rd$.
+In each setting, we derive upper bounds which imply that the PCR-LE estimate $\wh{f}$, and a test using the statistic $T = \|\wh{f}\|_2^2$, are statistically optimal methods for two classical problems in nonparametric regression: estimation and goodness-of-fit testing.
+
+At first glance, the optimality of PCR-LE is surprising.
+As a function of the sample size $n$, known rates of convergence of LE are much slower than the minimax-optimal rates of convergence over $H^s(\mc{X})$ (roughly speaking $n^{-2s/(2s + d)}$ for estimation and $n^{-4s/(4s + d)}$ for testing).
+This is not due to suboptimal theory for LE.
+Rather, it reflects a fundamental fact: for this problem, regression using learned features does not rely on accurately learning the features.
+It is essential to employ modes of analysis which exploit this fact in order to derive sharp rates of convergence for PCR-LE.
+We explain this in more detail after summarizing our main results.
+
+\paragraph{Sobolev spaces and spectral series regression.}
+To analyze PCR-LE, we work in a classical situation where the regression function is assumed to belong to a (Hilbert-)Sobolev space.
+For an open domain $\mc{X} \subseteq \Rd$, the Sobolev space $H^s(\mc{X})$ consists of all functions $f \in L^2(\mc{X})$ which are $s$-times weakly differentiable, with all order-$s$ partial derivatives $D^{\alpha}f \in L^2(\mc{X})$.
+We study regression over Sobolev spaces in part because, generally speaking, the minimax rates are well-understood; as mentioned before, when the domain $\mc{X}$ is full-dimensional they are $n^{-2s/(2s + d)}$ for estimation, and $n^{-4s/(4s + d)}$ for testing.
+For this reason, regression over Sobolev spaces is a good setting in which to see whether PCR-LE measures up to simpler and more classical minimax-optimal approaches, which have strong theoretical guarantees but are less often used in practice.
+
+Moreover, we view PCR-LE as being particularly well-suited for regression over Sobolev spaces due to their close connection with \emph{spectral series regression}.
+Spectral series regression computes empirical Fourier coefficients and truncates to the $K$-lowest frequency eigenfunctions, producing the estimate
+\begin{equation}
+\label{eqn:population-level_spectral_series}
+\wt{a}_k = \frac{1}{n}\sum_{i = 1}^{n} Y_i \psi_k(X_i), \quad \wt{f}(x) = \sum_{k = 1}^{K} \wt{a}_k \psi_k(x).
+\end{equation} 
+Under appropriate boundary conditions, the Sobolev spaces $H^s(\mc{X})$ can roughly be characterized as consisting of those functions $f = \sum_{k} a_k \psi_k \in L^2(\mc{X})$ for which the generalized Fourier coefficients $\{a_k\}_{k = 1}^{\infty}$ satisfy the decay condition $\sum_{k} a_k \rho_k^2 < \infty$ (See Section~\ref{subsec:spectral_projection} for more details).
+This decay condition justifies the truncated series estimator \eqref{eqn:population-level_spectral_series}, since it means the truncation will incur only a limited amount of bias for any $f_0 \in H^s(\mc{X})$.
+For this reason spectral series regression over Sobolev spaces has been well-studied since at least~\citet{rice1984}\footnote{And of course proposed much earlier in the context of density estimation by \citet{cencov1962}.}, and its optimality properties are by this point well-understood.
+
+PCR-LE serves as an empirical approximation to spectral series regression, since the eigenvectors $v_k$ are empirical approximations to the eigenfunctions $\psi_k$.
+Viewed in this light, a major advantage of PCR-LE is that it operates independently of the design distribution $P$.
+In the contrast, the spectral series estimator defined in~\eqref{eqn:population-level_spectral_series} relies on diagonalizing the density-weighted Laplacian $\Delta_P$.
+In our context, where $P$ is treated as unknown,~\eqref{eqn:population-level_spectral_series} must be viewed as an oracle method, and to emphasize this  we henceforth refer to~\eqref{eqn:population-level_spectral_series} as \emph{population-level spectral series regression}.
+On the other hand, naturally PCR-LE incurs some extra error by using an empirical approximation to the underlying basis $\{\psi_k\}_{k = 1}^{\infty}$, and our work shows that in many cases, this extra error is not enough to change the overall rate of convergence.
+
+\subsection{Main contributions}
+Summarized succinctly, our main contribution is to theoretically analyze nonparametric regression with PCR-LE and establish upper bounds which imply that this method often achieves optimal rates of convergence over Sobolev spaces.
+
+\paragraph{Rates of convergence: population-level spectral series regression.}
+As we have already mentioned, the minimax-optimal rates over Sobolev spaces are generally well-known, as are upper bounds for population-level spectral series methods which match these rates.
+However, we could not find precisely stated results applying to our setting, which is somewhat unusual in the following respects.
+\begin{enumerate}
+	\item We consider Sobolev spaces $H^s(\mc{X})$ for all combinations of $s$ and $d$.
+This includes the subcritical regime where the smoothness parameter $s$ satisfies $s < d/2$ and so $H^s(\mc{X})$ does not continuously embed into the space of continuous functions $C^0(\mc{X})$.
+	\item We consider general design distributions $P$, which may satisfy certain regularity conditions but are not limited to being, say, the uniform distribution over $[0,1]^d$. 
+\end{enumerate}
+For completeness, we analyze population-level spectral series methods in this general setting, and establish upper bounds showing that such methods converge at the ``usual'' rates of $n^{-2s/(2s + d)}$ for estimation and $n^{-4s/(4s + d)}$ for testing.
+This analysis relies heavily on certain asymptotic properties of the continuum eigenfunctions $\psi_k$ and eigenvalues $\rho_k$, which hold for quite general second-order differential operators $\mc{L}$ including the density-weighted Laplacian $\mc{L} = \Delta_P$.
+
+\paragraph{Rates of convergence: PCR-LE.}
+The rest of our results consist of various upper bounds on the rates of convergence for the PCR-LE estimator $\wh{f}$, and a test using the statistic $\wh{T} = \|\wh{f}\|_2^2$.
+These upper bounds quantify two important properties of PCR-LE: first, that it can take advantage of smooth higher-order derivatives, and second that it can adapt to low intrinsic dimension of the design distribution, each in an optimal manner.
+We consider two models for the design distribution $P$, the flat Euclidean and manifold models (See Section~\ref{subsec:regression_laplacian_eigenmaps} for the formal definitions of these models).
+In the first model, the design distribution $P$ has support $\mc{X}$ which is a full-dimensional set in $\Rd$.
+In this case, our main contributions are as follows:
 \begin{itemize}
-	\item Over the Sobolev space $H^{s}(\mc{X})$, we establish that the Laplacian eigenmaps estimator $\wh{f}$ has in-sample mean-squared error of at most on the order of $n^{-2s/(2s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d$.
-A test based on the statistic $\|\wh{f}\|_n^2$ has a squared critical radius on the order of $n^{-4s/(4s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d \in \{1,2,3,4\}$.
-	\item We also consider the behavior of Laplacian eigenmaps when the data satisfies a \emph{manifold hypothesis}, meaning the design points $\{X_1,\ldots,X_n\}$ belong to some low-dimensional submanifold of $\Rd$.
-In this case, it is known that the minimax rates over Sobolev spaces depend on the intrinsic dimension $m$ of the manifold: explicitly, the estimation rate is $n^{-2s/(2s + m)}$ and the testing rate is $n^{-4s/(4s + m)}$.
-We give upper bounds showing that Laplacian eigenmaps achieves these rates so long as $s \in \{1,2,3\}$.
-One can interpret these results as precisely quantifying a way in which Laplacian eigenmaps leverages structure, by adapting to the intrinsic dimension of the design.
-	\item The estimator $\wh{f}$ is defined only in-sample.
-We propose a simple out-of-sample extension based on kernel smoothing, and show that this extension adds negligible error, resulting in an estimator which is defined on all of $\mc{X}$ and has out-of-sample mean-squared error also on the order of $n^{-2s/(2s + d)}$.
+	\item  Over a ball in the Sobolev space $H^{s}(\mc{X})$, we establish that the PCR-LE estimator $\wh{f}$ has in-sample mean-squared error of at most on the order of $n^{-2s/(2s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d$.
+	\item We show that a test based on the statistic $\|\wh{f}\|_2^2$ has a squared critical radius on the order of $n^{-4s/(4s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d \in \{1,2,3,4\}$.
 \end{itemize}
-In all cases, our bounds also depend optimally on the radius $M$ of the Sobolev ball under consideration.
-
-To derive these results, we develop a framework for analysis reminiscent of that  classically used to analyze population-level spectral projection estimators, but ultimately quite different.
-The chief difference is in the approximation error (bias).
-For our purposes we must establish that with high probability, the eigenvectors of graph Laplacians (which are data-dependent and hence random objects) can be used to effectively approximate continuum Sobolev functions, whereas in the classical analysis an upper bound on the approximation error follows immediately from the aforementioned spectral characterization of Sobolev spaces.
-To tightly upper bound the approximation error of Laplacian eigenmaps, we rely on several recent results regarding concentration of the spectra of graph Laplacians around their population limits, and prove some additional results of this type which may be of independent interest.
-
-For some values of $s$ (number of derivatives) and $d$ (dimension), there do exist gaps between our upper bounds on the error of Laplacian eigenmaps and the minimax rates.
+We then consider the behavior of PCR-LE when the data satisfies a \emph{manifold hypothesis}, meaning the design distribution is supported on an (unknown) domain $\mc{X}$ which is a submanifold of $\Rd$ of intrinsic dimension $m \in \mathbb{N}, m < d$.
+In this case, our main contributions are as follows:
+\begin{itemize}
+	\item Over a ball in the Sobolev space $H^{s}(\mc{X})$, the PCR-LE estimator has in-sample mean squared error of at most $n^{-2s/(2s + m)}$, when $s \in \{1,2,3\}$ and for any $m \in \mathbb{N}$. 
+	\item A test based on the statistic $\|\wh{f}\|_2^2$ has a squared critical radius on the order of $n^{-4s/(4s + m)}$, when $s \in \{1,2,3\}$ and $m \in \{1,2,3,4\}$.
+\end{itemize}
+To the best of our knowledge, the minimax rates for nonparametric regression with random design over unknown manifolds have only been worked out for H\"{o}lder classes, and even in this case the calculations are only for $s \leq 2$ bounded derivatives~\citep{bickel2007,yang2016}.
+Our upper bounds confirm that these rates are the same for Sobolev spaces---at least when loss is measured in empirical norm---for the values of $s$ and $m$ mentioned above.
+
+In all these cases, our bounds also depend optimally on the radius $M$ of the Sobolev ball under consideration.
+However, for some values of $s$ (number of derivatives) and $d$ (dimension), there do exist gaps between our upper bounds on the error of PCR-LE and the minimax rates.
 Although we do not give corresponding lower bounds verifying the tightness of our analysis, we believe these gaps reflect the true behavior of the method rather than some looseness in our analysis, and we comment more on this at relevant parts in the text.
@@
 	\end{center}
-	\caption{Summary of Laplacian eigenmaps estimation rates over Sobolev balls.
+	\caption{Summary of PCR-LE estimation rates over Sobolev balls.
 Bold font marks minimax optimal rates.
 In each case, rates hold for all $d \in \mathbb{N}$ (under Model~\ref{def:model_flat_euclidean}), and for all $m \in \mathbb{N}, 1 < m < d$ (under Model~\ref{def:model_manifold}).
-Although we suppress it for simplicity, in all cases when the Laplacian eigenmaps estimator is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal}
+Although we suppress it for simplicity, in all cases when the PCR-LE estimator is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal.}
 	\label{tbl:estimation_rates}
@@
 	\end{center}
-	\caption{Summary of Laplacian eigenmaps testing rates over Sobolev balls.
+	\caption{Summary of PCR-LE testing rates over Sobolev balls.
 Bold font marks minimax optimal rates.
 Rates when $d > 4s$ assume that $f_0 \in L^4(\mc{X})$, and depend on $\|f_0\|_{L^4(\mc{X})}$.
-Although we suppress it for simplicity, in all cases when othe Laplacian eigenmaps test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal.}
+Although we suppress it for simplicity, in all cases when othe PCR-LE test is optimal, the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal.}
 	\label{tbl:testing_rates}
@@
 
-\paragraph{Related work.}
+\paragraph{Technical contributions.}
+One of our primary technical contributions is a strategy for analyzing regression methods using learned features which leads to sharp rates of convergence.
+
+At an abstract level PCR-LE is a two-stage algorithm: the first stage (LE) uses the observed design points $\{X_1,\ldots,X_n\}$ to learn a feature representation ($X_i \mapsto (v_{1,i},\ldots,v_{K,i})$) which is an empirical approximation to some ideal population-level representation ($X_i \mapsto (\psi_{1}(X_i),\ldots,\psi_{K}(X_i))$);  the second stage then applies a simple method (OLS) to the learned features, and obtains an estimate.
+At first blush, one might expect the error for PCR-LE to be likewise decomposed into two parts: first, the error with which the features are learned (i.e.
+error due to sampling variation in $X$), second, the error with which, given ideal population-level features, the regression function is learned (i.e.
+error due to noise in $Y$.)
+
+Crucially, our analysis \emph{does not} work in this way.
+This is important because, as already mentioned, all known upper bounds on the error with which the learned features (Laplacian eigenvectors) approximate their population-level limits (continuum Laplacian eigenfunctions) are much larger than the minimax rates of convergence over $H^s(\mc{X})$.
+For example,
+\begin{itemize}
+	\item The best known rates of convergence for graph Laplacian eigenvectors (due to~\cite{cheng2021}) are
+	\begin{equation}
+	\label{eqn:eigenvector_convergence}
+	\max_{1 \leq k \leq K}\|\sqrt{n}v_k - \psi_k\|_n^2 \leq C_{K} n^{-2/(4 + d)},
+	\end{equation}
+	where the constant $C_K$ depends on $K$ in an unknown manner.
+	\item The dependence of $C_K$ on $K$ is relevant for our interests, since in order to optimally trade off bias and variance we will need to choose $K$ is growing with $n$.
+The earlier works of \cite{burago2014,trillos2019} give upper bounds which, although implying weaker rates of convergence as a function of $n$, do keep track of $C_K$, and suggest that it should depend on the inverse of the spectral gap
+	\begin{equation}
+	\label{eqn:spectral_gap}
+	C_K = \frac{C}{\rho_{K + 1} - \rho_K} \geq C K^{1 - 2/d},
+	\end{equation}
+	where $C$ is a constant depending only on the dimension $d$, and the latter inequality follows whenever the eigenvalues $\rho_K$ obey the Weyl's Law asymptotic scaling $\rho_K \asymp K^{2/d}$.
+Thus, for a given $K(n)$ growing with $n$, when $d \geq 3$, the rates of convergence of $v_{K(n)} - \psi_{K(n)}$ are even slower than $n^{-2/(4 + d)}$.
+\end{itemize}
+Although these upper bounds may not reflect the true rate of convergence of graph Laplacian eigenvectors---this is still an active area of research, and no lower bounds are known---it seems very unlikely that the true rate matches the minimax estimation rate $n^{-2s/(2s + d)}$, which after all approaches the dimension-free rate $1/n$ for large values of $s$.
+
+Instead of relying on convergence of eigenvectors to eigenfunctions, our analysis proceeds via a bias-variance decomposition at the level of the graph. Focusing for simplicity on estimation, this is 
+\begin{equation}
+\label{eqn:bias_variance_estimation}
+\|f_0 - \wh{f}\|_n^2 \leq 2\bigl(\|f_0 - V_KV_K^{\top}f_0\|_n^2 + \|V_KV_K^{\top}({\bf Y} - f_0)\|_n^2\bigr) 
+\end{equation}
+The second term in~\eqref{eqn:bias_variance_estimation} is the variance, and as usual for OLS estimates depends only on the degrees of freedom $\mathrm{df}(\wh{f}) = \mathrm{tr}(V_KV_K^{\top}) = K$.
+More surprisingly, the first term (squared-bias) can also be upper bounded without appealing to the eigenfunctions $\psi_k$.
+Letting $S = L^{\dagger}$ be the pseudo-inverse of the graph Laplacian, a few lines of algebra give
+\begin{equation}
+\label{eqn:bias_estimation}
+\begin{aligned}
+\|(I - V_KV_K^{\top})f_0\|_n^2 & = \|(I - V_KV_K^{\top})S^{s/2}L^{s/2}f_0\|_n^2 \leq \|(I - V_KV_K^{\top})S^{s/2}\|_{op}^2 \|L^{s/2}f_0\|_n^2 = \frac{f_0^{\top} L^s f_0}{n \lambda_{K + 1}^{s}}.
+\end{aligned}
+\end{equation}
+Thus we obtain an upper bound on the in-sample sample mean squared error $\|\wh{f} - f_0\|_n^2$ that is \emph{independent of the error with which graph Laplacian eigenvectors $v_k$ approximate population-level eigenfunctions $\psi_k$}.
+
+Instead, our upper bound on the error of PCR-LE is determined by a pair of graph functionals: the quadratic form $f_0^{\top}L^s f_0$, and the graph Laplacian eigenvalue $\lambda_{K + 1}$.
+Theoretically speaking, this brings two advantages over directly analyzing convergence of eigenvectors to eigenfunctions.
+The first is that sharper rates  of convergence are known for these graph functionals than for the eigenvectors $v_k$.
+The second, arguably more important, is that in order to obtain upper bounds on $\|\wh{f} - f_0\|_n^2$ we do not require that these graph functionals themselves converge to population-level limits, but only that they be stochastically bounded on the right order.
+To derive our ultimate upper bounds on the error of PCR-LE we use some existing results regarding neighborhood graph Laplacians, and prove some new ones which may be of independent interest.
+
+A last observation regarding the difference between feature learning and regression using learned features.
+To balance the bias and variance terms in~\eqref{eqn:bias_variance_estimation}, we will end up using $K(n) := n^{d/(2s + d)}$ eigenvectors.
+Although this is the usual choice for spectral series estimation over order-$s$ Sobolev spaces, in the context of PCR-LE it leads to a striking conclusion.
+Namely, examining~\eqref{eqn:eigenvector_convergence} and~\eqref{eqn:spectral_gap}, we see that for all dimensions $d \geq 3$, known upper bounds on $\|\sqrt{n} v_{K(n)} - \psi_{K(n)}\|_n^2$ do not converge to $0$ as $n \to \infty$.
+In words, it is not merely the case that PCR-LE converges at a faster rate than LE itself, but that PCR-LE can profitably use eigenvectors for regression which may be inconsistent estimates of their population-level limits.
+
+To summarize, our work demonstrates, broadly speaking, that regression using learned features can be analyzed independent of the accuracy of the features themselves.
+Regression using learned features is a general and widely applied paradigm, and we believe this observation may have consequences outside of its application to PCR-LE in this work.
+
+\subsection{Related work}
+
+Much of the work regarding regression using neighborhood graph Laplacians deals with \emph{semi-supervised learning}, where in addition to the labeled data $(X_1,Y_1),\ldots,(X_n,Y_n)$ one observes unlabeled points $(X_{n + 1},\ldots,X_{N})$, and the task is to produce an estimate at labeled and unlabeled points alike.
+To this end, the landmark paper of \cite{zhu2003semisupervised} proposed to interpolate the observed values by~\emph{harmonic extension}, i.e.
+compute the Laplacian matrix $L_N$  corresponding to a graph formed over all design points $X_1,\ldots,X_N$, and then solve the constrained problem
+\begin{equation*}
+\minimize_{f \in \Reals^N} f^{\top} L_N f \quad \mathrm{subject\,\,to}~~~ f_i = Y_i~~\textrm{for $i = 1,\ldots,n$.}
+\end{equation*}
+Conventional wisdom says that harmonic extension is sensible only when the responses are noiseless, $Y_i = f_0(X_i)$, and that in the noisy setting one should instead solve the penalized formulation
+\begin{equation}
+\label{eqn:graph_laplacian_regularization_ssl}
+\wc{f} := \argmin_{f \in \Reals^N} \|{\bf Y} - f\|_n^2 + \lambda f^{\top} L_N f.
+\end{equation}
+Notwithstanding their intuitive appeal, both the constrained and penalized problems have issues when $d > 1$ and $n/N \to 0$: the estimates tend towards degeneracy, meaning they are ``spiky'' at labeled data points and close to constant everywhere else~\citep{nadler09,calder2019b, calder2020}.
+One solution to this problem is to instead use Laplacian Eigenmaps for semi-supervised learning (SSL-LE), i.e.
+compute the eigendecomposition $L_N = \sum_{k = 1}^{N} \lambda_k u_k u_k^{\top}$ and solve the problem
+\begin{equation}
+\label{eqn:laplacian_eigenmaps_ssl}
+\wb{f} := \argmin_{f \in \mathrm{span}\{u_1,\ldots,u_K\}} \|{\bf Y} - f\|_n^2.
+\end{equation}
+\cite{zhou2011,lee2016} analyze SSL-LE in a particular asymptotic regime where the number of labeled points $n$ is held fixed while the number of unlabeled points $N - n \to \infty$.
+They show that the estimator $\wb{f}$ achieves minimax-optimal rates of convergence over $H^s(\mc{X})$ as a function of $n$.
+Of course, as $N - n \to \infty$ the low-frequency eigenvectors of the graph Laplacian $L_N$ all converge to their continuum limits.
+This means in particular that for $K(n) = n^{d/(2s + d)}$,
+\begin{equation*}
+\lim_{N \to \infty} \max_{1 \leq K \leq K(n)} \|\sqrt{N} u_k - \psi_k\|_N^2 = 0,
+\end{equation*}
+and consequently when $\wb{f}$ is optimally tuned, $\lim_{N \to \infty} \|\wb{f} - \wt{f}\|_N^2 = 0$.
+Thus the analysis of SSL-LE reduces to that of population-level spectral series regression.
+The supervised setting (where $N = n$) we consider in this work is very different, and analyzing PCR-LE necessitates an entirely different approach, as explained in the preceding section.
+
+There has been much less work, relatively speaking, regarding regression with neighborhood graph Laplacians in the supervised setting, which is the focus of this work.
+\citet{lee2016} also analyze a variant of PCR-LE, but derive suboptimal rates of convergence.
+\citet{trillos2020} study nonparametric regression via graph Laplacian regularization, where the estimator is computed by solving~\eqref{eqn:graph_laplacian_regularization_ssl}, but with no unlabeled data.
+They establish the uniform upper bound $\max_{i = 1,\ldots,n}|\wc{f}(X_i) - f_0(X_i)| \leq C n^{-2/(2 + d)}$ under the assumption $f_0 \in C^2(\mc{X})$, which is slower than the minimax rate $n^{-4/(4 + d)}$ for this class.
+In a previous paper~\citep{green2021}, we (the authors) also considered nonparametric regression via graph Laplacian regularization, when the regression function $f_0 \in H^1(\mc{X})$ and loss is measured using in-sample mean squared error.
+We found that graph Laplacian regularization is consistent as $n \to \infty$, and optimal (using mean-squared error) over the first-order Sobolev space $H^1(\mc{X})$ when $d \in \{1,2,3,4\}$.
+However, graph Laplacian regularization neither takes advantage of smooth higher-order derivatives, nor is it provably optimal over $H^1(\mc{X})$ for all dimensions $d$.
+One of our motivations for considering PCR-LE was to find an estimator which addressed these deficiencies.
+In this work we indeed find that PCR-LE has much stronger optimality properties than those known for graph Laplacian regularization, or indeed any other method for regression using neighborhood graphs.
 
@@
 
-In the random design setting the analysis of graph-based approaches to supervised learning is more challenging, and only a few works have studied the matter.
-\cite{zhou2011} provide an asymptotic analysis of the Laplacian eigenmaps estimator in the semi-supervised setting, as the number of unlabeled samples tends to infinity; in this case the estimator is equivalent to population-level spectral projection, and there is no extra error incurred by using data-dependent eigenvectors.
-\cite{lee2016} consider the \emph{diffusion maps} estimator---which uses the eigenvectors of a differently normalized graph Laplacian---in a similar supervised setting to our own, but prove suboptimal rates of convergence.
-
-Finally, \citet{trillos2020,green2021} study \emph{graph Laplacian regularization}, a graph-based penalized least squares method which uses a quadratic form involving the graph Laplacian as a penalty.
-Laplacian regularization can, like Laplacian eigenmaps, be cast as a spectral method.
-The difference is that Laplacian regularization induces some smooth shrinkage based on the eigenvalue, whereas Laplacian eigenmaps corresponds to a hard spectral cutoff.
-Interestingly, this difference leads to a stark contrast in the ultimate statistical properties of the two approaches.
-For instance, in~\cite{green2021} Laplacian regularization is shown to be optimal only for the first-order Sobolev space $H^1(\mc{X})$, and even then only when $d \in \{1,2,3,4\}$.
-In contrast, in this work we show that the Laplacian eigenmaps estimator is optimal for all $s$ and $d$.
-
 \paragraph{Roadmap.}
 We now outline the structure of the rest of this paper.
-In Section~\ref{sec:setup_main_results}, we give our formal modeling assumptions, and precisely define the Laplacian eigenmaps methods we study.
-Propositions~\ref{prop:spectral_series_estimation} and ~\ref{prop:spectral_series_testing}, in Section~\ref{subsec:spectral_projection}, show that under rather general (nonparametric) conditions on the design distribution, population-level spectral projection methods achieve minimax rates of convergence over Sobolev classes.
-Then in Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps},~\ref{sec:manifold_adaptivity}, and \ref{sec:out_of_sample}, we give our main upper bounds on the error of Laplacian eigenmaps (summarized above); these upper bounds hold under similarly general conditions, and imply that Laplacian eigenmaps estimators and tests are also minimax rate-optimal.
-In Section~\ref{sec:experiments} we examine the empirical behavior of Laplacian eigenmaps, and show that even at moderate sample sizes Laplacian eigenmaps is competitive with population-level spectral projection.
+In Section~\ref{sec:setup_main_results}, we give our formal modeling assumptions, and precisely define the PCR-LE estimator and test we study.
+Propositions~\ref{prop:spectral_series_estimation} and ~\ref{prop:spectral_series_testing}, in Section~\ref{subsec:spectral_projection}, show that under rather general (nonparametric) conditions on the design distribution, population-level spectral series methods achieve minimax rates of convergence over Sobolev classes.
+Then in Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and ~\ref{sec:manifold_adaptivity} we give our main upper bounds on the error of PCR-LE.
+These upper bounds (summarized above) hold under similarly general conditions, and imply that the PCR-LE estimator and test are also minimax rate-optimal.
+In Section~\ref{sec:experiments} we examine the empirical behavior of PCR-LE, and show that even at moderate sample sizes PCR-LE is competitive with population-level spectral series regression.
 We conclude with some discussion in Section~\ref{sec:discussion}.
@@
 
-We begin in Sections~\ref{subsec:regression_laplacian_eigenmaps}-\ref{subsec:laplacian_eigenmaps} by precisely defining the models (random design points, Sobolev regression functions) and methods (Laplacian eigenmaps) under consideration.
-Then in Section~\ref{subsec:spectral_projection}, we connect Sobolev spaces to Laplacian eigenmaps using the spectrum of a density-dependent Laplacian operator, and show in Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} that projection methods which use the eigenfunctions of this operator are statistically optimal.
+We begin in Sections~\ref{subsec:regression_laplacian_eigenmaps}-\ref{subsec:laplacian_eigenmaps} by precisely defining the models (random design points, Sobolev regression functions) and methods (Principal Components Regression with Laplacian Eigenmaps) under consideration.
+Then in Section~\ref{subsec:spectral_projection}, we analyze the behavior of population-level spectral series methods.
 
@@
 
-We will always operate in the usual setting of nonparametric regression with random design.
-We observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$; the design points $X_1,\ldots,X_n$ are sampled from a distribution $P$ with support $\mc{X} \subseteq \Rd$, and the responses follow the signal plus noise model
-\begin{equation}
-\label{eqn:model}
-Y_i = f_0(X_i) + w_i,
-\end{equation}
-with regression function $f_0: \mc{X} \to \Reals$, and $w_i \sim N(0,1)$ independent Gaussian noise.
-% (AG 8/4/21): Removed the comment regarding noise with variance = \sigma^2.
-
-We now formulate two models, which differ in the assumed nature of the support $\mc{X}$ of the design distribution $P$: the \emph{flat Euclidean} and \emph{manifold} models.
+As mentioned, we will always operate in the usual setting of nonparametric regression with random design.
+We observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$, where the design points $X_1,\ldots,X_n$ are sampled from a distribution $P$ with support $\mc{X} \subseteq \Rd$, and the responses follow~\eqref{eqn:model}.
+We now formulate two models for the design distribution $P$ and regression function $f_0$: the \emph{flat Euclidean} and \emph{manifold} models.
 
@@
 
-\subsection{Laplacian Eigenmaps}
+\subsection{Principal Components Regression with Laplacian Eigenmaps (PCR-LE)}
 \label{subsec:laplacian_eigenmaps}
@@
 
-The Laplacian eigenmaps estimator $\wh{f}$ simply projects the response vector ${\bf Y} = (Y_1,\ldots,Y_n)$ onto the first $K$ eigenvectors of $L_{n,\varepsilon}$: letting $V_K \in \Reals^{n \times K}$ be the matrix with columns $V_{K,k} = v_k$, we have that
+The PCR-LE estimator $\wh{f}$ defined in~\eqref{eqn:pcr-le} simply projects the response vector ${\bf Y}$ onto the first $K$ eigenvectors of $L_{n,\varepsilon}$.
+Since the eigenvectors of the graph Laplacian are orthonormal with respect to the Euclidean inner product on $\Reals^n$, we can more simply write this as
 \begin{equation}
 \label{eqn:laplacian_eigenmaps_estimator}
-\wh{f} := \sum_{k = 1}^{K} \dotp{{\bf Y}}{v_k}_{2} v_k = V_K V_K^{\top} {\bf Y}.
-\end{equation} 
-If $\wh{f}$ is a reasonable estimate of $f_0$, then the Laplacian eigenmaps test statistic,
+\wh{f} = V_K V_K^{\top} {\bf Y},
+\end{equation} 
+where $V_K \in \Reals^{n \times K}$ is the matrix with $k$th column $V_{K,k} = v_k$. The PCR-LE test statistic is 
 \begin{equation}
@@
 \end{equation}
-is in turn a reasonable estimate of $\|f_0\|_{P}^2$, and can be used in the \emph{signal detection} problem to distinguish whether or not $f_0 = 0$.
-
-\subsection{Spectral Projection using Continuum Eigenfunctions}
+and can be used in the \emph{signal detection} problem to distinguish whether or not $f_0 = 0$.
+
+\subsection{Spectral series regression over Sobolev spaces}
 \label{subsec:spectral_projection}
-Laplacian eigenmaps serves as a data-dependent alternative to more classical population-level spectral projection methods, which operate by projecting the response vector ${\bf Y}$ onto a subspace spanned by eigenfunctions of some differential operator.
-For instance, projecting ${\bf Y}$ onto the eigenfunctions of the Laplacian operator $\Delta = \sum_{i = 1}^{d} \partial^2f/\partial x_i^2$ corresponds to estimating $f_0$ using noisy empirical Fourier coefficients.
-
-We will focus on a density-dependent alternative to the Laplacian,  $\Delta_P:C^2(\mc{X}) \to L^2(\mc{X})$,
-\begin{equation}
-\label{eqn:laplace_beltrami}
-\Delta_Pf := -\frac{1}{p} \mathrm{div}(p^2\nabla f),
-\end{equation}
-which has enumerable eigenvalue/eigenfunction pairs $(\lambda_k(\Delta_P),\psi_k),(\lambda_2(\Delta_P),\psi_2),\ldots$ defined according to
+We now establish some upper bounds on the error of population-level spectral series regression when $f_0 \in H^s(\mc{X})$, which imply that such methods achieve optimal rates of convergence for both estimation and testing.
+The upper bounds we establish are ``usual'' in the sense that they match the rates $n^{-2s/(2s + d)}$ (estimation) and $n^{-4s/(4s + d)}$ (testing) which are already known in many cases.
+However, they are unusual in that we treat both the case where $s < d/2$ and thus the Sobolev space $H^s(\mc{X})$ does not continuously embed into $C(\mc{X})$, and the case where $P$ is not the uniform distribution over the unit cube.
+The upper bounds given in this section serve two purposes: first, to clarify what the rates are in these more unusual settings; second, to show that population-level spectral series regression can obtain the optimal rates in such settings.
+This latter point is important since the method we focus on for the most part, PCR-LE, is an empirical approximation to population-level spectral series regression.
+
+\paragraph{Spectrally defined Sobolev spaces.}
+Let $\mc{X}$ be an open domain which satisfies the conditions of Model~\ref{def:model_flat_euclidean}.
+Recalling the density-weighted Laplacian $\Delta$, defined in~\eqref{eqn:density-weighted-laplace}, we consider the eigenvector equation with Neumann boundary conditions,
 \begin{equation}
 \label{eqn:laplace_beltrami_eigenproblem}
-\Delta_P\psi = \lambda(\Delta_P) \psi, \quad \frac{\partial}{\partial{\bf n}}\psi = 0~~\textrm{on $\partial \mc{X}$,}
-\end{equation}
-and sorted as usual in ascending order of eigenvalue.\footnote{For formal justification of the fact that~\eqref{eqn:laplace_beltrami_eigenproblem} has a discrete spectrum under either Model~\ref{def:model_flat_euclidean} or~\ref{def:model_manifold}, see~\cite{garciatrillos18,trillos2019}.} Formally, the classical spectral series estimator and test statistic are\footnote{Technically speaking, $\wt{f}$ is not the projection of ${\bf Y}$ onto the subspace of $L^2(P_n)$ spanned by $\psi_1,\ldots,\psi_K$, since $\dotp{\psi_k}{\psi_{\ell}}_n \neq 0$.
-However, these eigenfunctions are very close to orthogonal in $L^2(P_n)$, orthogonalization will have a negligible impact on the overall statistical properties, and so for simplicity we consider the methods as defined in~\eqref{eqn:classical_spectral_projection}.}
-\begin{equation}
-\label{eqn:classical_spectral_projection}
-\wt{f}(x) := \sum_{k = 1}^{K} \dotp{{\bf Y}}{\psi_k}_{n} \psi_k(x),\quad\textrm{and}\quad \wt{T} := \|\wt{f}\|_P^2.
-\end{equation}
-
-We focus on the particular differential operator $\Delta_P$ because the spectrum of $\Delta_P$ is the population-level limit of the graph Laplacian spectrum: for any fixed $k \in \mathbb{N}$, as the number of samples $n \to \infty$, the graph radius $\varepsilon \to 0$, and $n\varepsilon^{2 + \dim(\mc{X})} \to \infty$,
-\begin{equation}
-\lambda_k \to \lambda_k(\Delta_P), \quad\textrm{and}\quad \Bigl\|\frac{1}{n}v_k - \psi_k\Bigr\|_n^2 \to 0.
-\end{equation}
-This holds in either the flat Euclidean (Model~\ref{def:model_flat_euclidean}) or manifold (Model~\ref{def:model_manifold}) setups~\citep{garciatrillos18,trillos2019}.
-It follows that for any fixed number of eigenvectors $K \in \mathbb{N}$, both $\wh{f} \to \wt{f}$ and $\wh{T} \to \wt{T}$, and in this way Laplacian eigenmaps can be formally tied to more classical spectral projection methods for regression.
-
-On the other hand, the eigenvalues and eigenfunctions of $\Delta_P$ can also be used to give a spectral definition of Sobolev spaces on general domains $\mc{X}$. Consider the ellipsoid
+\Delta_P\psi = \rho \psi, \quad \frac{\partial}{\partial{\bf n}}\psi = 0~~\textrm{on $\partial \mc{X}$.}
+\end{equation}
+Under Model~\ref{def:model_flat_euclidean}, the eigenvector equation~\eqref{eqn:laplace_beltrami_eigenproblem} has enumerable solutions $(\rho_1,\psi_1),(\rho_2,\psi_2),\ldots$, sorted as usual in ascending order of eigenvalue~\citep{garciatrillos18,trillos2019}.
+These eigenvalues and eigenfunctions can be used to give a spectral definition of Sobolev spaces.
+Consider the ellipsoid
 \begin{equation}
 \label{eqn:sobolev_ellipsoid}
-\mc{H}^{s}(\mc{X}) := \Bigl\{\sum_{k = 1}^{\infty} a_k \psi_k \in L^2(\mc{X}):  \sum_{k = 1}^{\infty} a_k^2 \lambda_{k}^s(\Delta_P) \leq M^2 \Bigr\},
-\end{equation}
-equipped with the norm $\|\sum_{k = 1}^{\infty} a_k \psi_k\|_{\mc{H}^s(\mc{X})}^2 = \sum_{k = 1}^{\infty} a_k^2 \lambda_{k}^s(\Delta_P)$.
+\mc{H}^{s}(\mc{X}) := \Bigl\{\sum_{k = 1}^{\infty} a_k \psi_k \in L^2(\mc{X}):  \sum_{k = 1}^{\infty} a_k^2 \rho_k^s \leq M^2 \Bigr\},
+\end{equation}
+equipped with the norm $\|\sum_{k = 1}^{\infty} a_k \psi_k\|_{\mc{H}^s(\mc{X})}^2 = \sum_{k = 1}^{\infty} a_k^2 \rho_k^s$.
 Under appropriate regularity conditions $\mc{H}^s(\mc{X})$ consists of functions $f \in H^s(\mc{X})$ which also satisfy some additional boundary conditions.
@@
 \end{equation}
-and likewise $\mc{H}^{2s + 1}(\mc{X}) = \mc{H}^{2s}(\mc{X}) \cap H^{2s + 1}(\mc{X})$ for any $s \geq 0$; additionally, the norms $\|\cdot\|_{\mc{H}^s(\mc{X})}$ and $\|\cdot\|_{H^s(\mc{X})}$ are equivalent.
-
-This spectral formulation suggests that population-level spectral projection methods---and in turn Laplacian eigenmaps---are natural choices for regression over Sobolev spaces.
-Furthermore, since these population-level methods serve, in a sense, as the infinite-data limit of Laplacian eigenmaps, the statistical properties of the former---which are generally speaking better understood and easier to derive---should intuitively shed some light on the properties of the latter.
-To this end, we now state a pair of results, Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, which establish that population-level spectral projection methods are statistically optimal for nonparametric estimation and testing over Sobolev ellipsoids.
+and likewise $\mc{H}^{2s + 1}(\mc{X}) = \mc{H}^{2s}(\mc{X}) \cap H^{2s + 1}(\mc{X})$ for any $s \geq 0$.
+Additionally, the norms $\|\cdot\|_{\mc{H}^s(\mc{X})}$ and $\|\cdot\|_{H^s(\mc{X})}$ are equivalent.
+
+% (AG 10/16/21): The following paragraph seemed overkill. Commented out for now in case it proves useful later.
+% This spectral formulation suggests that population-level spectral series methods---and in turn PCR-LE---are natural choices for regression over Sobolev spaces.
+Furthermore, since these population-level methods serve, in a sense, as the infinite-data limit of PCR-LE, the statistical properties of the former---which are generally speaking better understood and easier to derive---should intuitively shed some light on the properties of the latter.
+To this end, we now state a pair of results, Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, which establish that population-level spectral series methods are statistically optimal for nonparametric estimation and testing over Sobolev ellipsoids.
 These results are generalizations of previously known upper bounds on the estimation and testing error of classical spectral methods~\citep{tsybakov08,ingster2009}, but they hold under less stringent (nonparametric) conditions on the design distribution $P$.
-This latter point is important, since as we have argued one attractive quality of Laplacian eigenmaps is its adaptivity to general design distributions.
+This latter point is important, since as we have argued one attractive quality of PCR-LE is its adaptivity to general design distributions.
 
 \paragraph{Estimation.}
-We begin with an upper bound on the $L^2(P)$ risk of $\wt{f}$.
+Recall the population-level spectral series estimator $\wt{f}$ defined in~\eqref{eqn:population-level_spectral_series}. We now give an upper bound on the $L^2(P)$ risk of $\wt{f}$.
 \begin{proposition}
 	\label{prop:spectral_series_estimation}
-	Suppose data is observed according to Model~\ref{def:model_flat_euclidean}, and that the density $p$ is known.
-Suppose additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$, $f_0 \in \mc{H}^{s}(\mc{X};M)$ and $\|f_0\|_P^2 \leq 1$.
-Then there exists a constant $C$ which does not depend on $f_0$, $M$ or $n$ such that the following statement holds: if the spectral projection estimator $\wt{f}$ is computed with parameter $K = \max\{\floor{M^2n}^{d/(2s + d)},1\}$, then
+	Suppose data is observed according to Model~\ref{def:model_flat_euclidean}.
+Assume additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$, $f_0 \in \mc{H}^{s}(\mc{X};M)$ and $\|f_0\|_P^2 \leq 1$.
+Then there exists a constant $C$ which does not depend on $f_0$, $M$ or $n$ such that the following statement holds: if the population-level spectral series estimator $\wt{f}$ is computed with parameter $K = \max\{\floor{M^2n}^{d/(2s + d)},1\}$, then
 	\begin{equation}
@@
 \end{proposition}
-When the Sobolev ball radius $n^{-1/2} \lesssim M$, the upper bound in~\eqref{eqn:spectral_series_estimation} is on the order of $M^2(M^2n)^{-2s/(2s + d)}$, which is the well-known minimax rate for estimation over Sobolev classes (see e.g.~\cite{gyorfi2006,wasserman2006,tsybakov08} and references therein, and specifically Theorem~3.2 of~\cite{gyorfi2006} for a matching lower bound in the context of nonparametric regression with random design).
-
-% (AG 8/19/21): I would like to strengthen this statement to not require $M \asymp 1$, but I do not know where I can find the rate M^2(M^2n)^{-2s/(2s + d)} for all s and d combinations.
-In fact, I don't even know of a paper that states the M^2(M^2n)^{-2s/(2s + d)} rate for L^2(P) risk in random design.
-
-We now give the proof of Proposition~\ref{prop:spectral_series_estimation}.
-The structure of the analysis, which is fairly classical and straightforward, can be usefully compared to our analysis of Laplacian eigenmaps (see Section~\ref{subsec:analysis}).
-
-\paragraph{Proof of Proposition~\ref{prop:spectral_series_estimation}.}
-We decompose risk into squared bias and variance,
-\begin{equation}
-\label{pf:spectral_series_estimation_0}
-\Ebb \|\wt{f} - f_0\|_P^2 = \Ebb\| \Ebb[\wt{f}]  - f_0\|_P^2 + \Ebb\| \wt{f} - \Ebb[\wt{f}]\|_P^2.
-\end{equation}
-Since the eigenfunctions $\{\psi_k\}$ form an orthonormal basis of $L^2(P)$, and $f_0 \in \mc{H}^{s}(\mc{X}) \subseteq L^2(P)$, we can write the squared bias in terms of squared Fourier coefficients of $f_0$, leading to the following upper bound,
-\begin{equation*}
-\|f_0 - \Ebb[\wt{f}]\|_P^2 = \sum_{k = K + 1}^{\infty}  \dotp{f_0}{\psi_k}_P^2 \leq  \frac{1}{\{\lambda_{K + 1}(\Delta_P)\}^s} \sum_{k = K + 1}^{\infty} \{\lambda_{k + 1}(\Delta_P)\}^s \dotp{f_0}{\psi_k}_P^2 \leq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}}{\{\lambda_{K + 1}(\Delta_P)\}^s}.
-\end{equation*}
-On the other hand, the variance term can be written as the sum of the variance of each empirical Fourier coefficient, and subsequently by the law of total variance we derive that
-\begin{align}
-\label{pf:spectral_series_estimation_2}
-\Ebb\| \wt{f} - \Ebb[\wt{f}]\|_P^2 = \sum_{k = 1}^{K} \Var\Bigl[\dotp{{\bf Y}}{\psi_k}_n\Bigr] & = \sum_{k = 1}^{K} \Var\Bigl[\Ebb[\dotp{Y}{\psi_k}_n|{\bf X}\Bigr] + \Ebb\Bigl[\Var[\dotp{Y}{\psi_k}_n|{\bf X}\Bigr] \nonumber \\
-& = \sum_{k = 1}^{K} \Var\Bigl[\dotp{f_0}{\psi_k}_n\Bigr] + \frac{1}{n}\Ebb\Bigl[\|\psi_k\|_n^2\Bigr] \nonumber \\
-& \leq \frac{K}{n} + \frac{1}{n}\sum_{k = 1}^{K}\Ebb\Bigl[\Bigl(f_0(X)\psi_k(X)\Bigr)^2\Bigr].
-\end{align}
-Consequently,
-\begin{equation}
-\label{pf:spectral_series_estimation_1}
-\Ebb \|\wt{f} - f_0\|_P^2 \leq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\bigl[\lambda_{K + 1}(\Delta_P)\bigr]^s} + \frac{K}{n} + \frac{1}{n}\Ebb\Bigl[(f_0(X))^2 \cdot \sum_{k = 1}^{K} (\psi_k(X))^2\Bigr].
-\end{equation}
-The claim of the proposition then follows from variants of two classical results in spectral geometry.
-The first is a Weyl's Law asymptotic scaling of the eigenvalues of $\Delta_P$ due to~\cite{dunlop2020}; formally, there exist constants $c$ and $C$ (which will depend on $P$ and $d$) such that
-\begin{equation}
-\label{eqn:weyl}
-ck^{2/d} \leq \lambda_k(\Delta_P) \leq Ck^{2/d}\quad\textrm{for all $k \in \mathbb{N}$, $k \geq 2$}.
-\end{equation}
-The second is a local analog to Weyl's Law, which says that there exists a constant $C$ (again depending on $P$ and $d$) such that
-\begin{equation}
-\label{eqn:local_weyl}
-\sup_{x \in \mc{X}}\biggl\{\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2\biggr\} \leq CK \quad\textrm{for all $K \in \mathbb{N}$}.
-\end{equation}
-Equation~\eqref{eqn:local_weyl} is a direct implication of~\eqref{eqn:weyl} along with Theorem 17.5.3 of~\cite{hormander1973}.
-Plugging the upper bounds~\eqref{eqn:weyl} and~\eqref{eqn:local_weyl} back into~\eqref{pf:spectral_series_estimation_1}, we conclude that
-\begin{equation}
-\label{pf:spectral_series_estimation_3}
-\Ebb \|\wt{f} - f_0\|_P^2 \leq C\biggl(\frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{(K + 1)^{2s/d}} + \frac{K}{n}\biggr).
-\end{equation}
-If $n^{-1/2} \geq M$, then taking $K = 1$ implies $\Ebb \|\wt{f} - f_0\|_P^2 \leq C(M^2 + 1/n)$.
-Otherwise, setting $K = \floor{M^2n}^{d/(2s + d)}$ balances squared bias and variance, and yields the claim.
-\qed.
+When the Sobolev ball radius $n^{-1/2} \lesssim M$, the upper bound in~\eqref{eqn:spectral_series_estimation} is on the order of $M^2(M^2n)^{-2s/(2s + d)}$.
+This is well known to be the minimax rate of estimation over the Sobolev classes $H^s([0,1]^d;M)$ when $s > d/2$; see e.g.~\cite{gyorfi2006,wasserman2006,tsybakov08} and references therein, and specifically Theorem~3.2 of~\cite{gyorfi2006} for a matching lower bound in the context of nonparametric regression with random design.
+On the other hand there seems to have been much less study of minimax rates over $H^s([0,1]^d;M)$ when $s < d/2$.
+In this \emph{subcritical} regime, the Sobolev space contains functions without continuous representatives, and certain questions become more subtle; see our remark after Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}.
+However, Proposition~\ref{prop:spectral_series_estimation} confirms that in this regime the minimax rates (with loss measured in squared-$L^2(P)$ norm) are still on the order of $M^2(M^2n)^{-2s/(2s + d)}$, since a matching lower bound follows from the known estimation rates over $C^s([0,1]^d) \subseteq H^s([0,1]^d)$~\citep{stone1980}.
 
@@
 
-For the signal detection problem, the population-level spectral projection test $\wt{\varphi} := \1\{\wt{T} \geq K/N + \sqrt{2K/an^2}\}$ has bounded Type I error, $\Ebb_{0}[\wt{\varphi}] \leq a (1 + o(1))$.
+For the signal detection problem, the population-level spectral series test $\wt{\varphi} := \1\{\wt{T} \geq K/N + \sqrt{2K/an^2}\}$ has bounded Type I error, $\Ebb_{0}[\wt{\varphi}] \leq a (1 + o(1))$.
 Proposition~\ref{prop:spectral_series_testing} gives an upper bound on the Type II error that holds over all $f_0 \in \mc{H}^s(\mc{X};M)$ for which $\|f_0\|_P^2$ is sufficiently large.
@@
 Suppose additionally that $\partial \mc{X} \in C^{1,1}$, $p \in C^{\infty}(\mc{X})$, $f_0 \in \mc{H}^{s}(\mc{X};M)$ for some $s > d/4$, and $\|f_0\|_{L^4(\mc{X})}^4 \leq 1$.
-Then there exists a constant $C$ which does not depend on $f_0$, $M$ or $n$ such that the following statement holds: if the spectral projection test $\wt{\varphi}$ is computed with parameter $K = \max\{\floor{M^2n}^{2d/(4s + d)},1\}$, and if
+Then there exists a constant $C$ which does not depend on $f_0$, $M$ or $n$ such that the following statement holds: if the population-level spectral series test $\wt{\varphi}$ is computed with parameter $K = \max\{\floor{M^2n}^{2d/(4s + d)},1\}$, and if
 	\begin{equation}
@@
 In other words, the regression testing problem over Sobolev spaces fundamentally changes when $s \leq d/4$, and hereafter when we discuss testing we will limit our consideration to $s > d/4$.
-\paragraph{Proof of Proposition~\ref{prop:spectral_series_testing}.}
-We briefly lay out the main ideas needed to prove Proposition~\ref{prop:spectral_series_testing}, following the lead of~\cite{ingster2009} who prove a similar result in the special case where $M = 1$ and $P$ is the uniform distribution over $\mc{X} = [0,1]^d$, and referring to that work for more details.
-
-We begin by computing the first two moments of the test statistic $\wt{T}$. The expectation is
-\begin{align*}
-\Ebb[\wt{T}] = \frac{(n - 1)}{n} \sum_{k = 1}^{K} \dotp{f_0}{\psi_K}_P^2 + \frac{K}{n} + \Ebb\Bigl[(f_0(X))^2 \dot \sum_{k = 1}^{K} (\psi_k(X))^2\Bigr],
-\end{align*}
-and from~\eqref{eqn:weyl} (Weyl's Law) we have that under the alternative $f_0 \neq 0$,
-\begin{equation*}
-\Ebb_{f_0}[\wt{T}] \geq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\lambda_{K + 1}^s(\Delta_{P})} + \frac{K}{n}.
-\end{equation*}
-To compute the variance, we decompose $\wt{T} = \wt{T}_{1,1} + \wt{T}_{1,2} + \wt{T}_{1,3} + \wt{T}_2$ into the sum of 3 U-statistics and the remaining diagonal terms, defined in terms of the equivalent kernel $\kappa(x,x') = \sum_{k = 1}^{K} \psi_k(x) \psi_k(x')$ as,
-\begin{align*}
-T_{1,1} & := \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} w_i w_j \kappa(X_i,X_j),\quad && T_{1,2} := \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} \bigl(w_i f_0(X_j) + w_jf_0(X_i)\bigr)\kappa(X_i,X_j) \\
-T_{1,3} & := \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} f_0(X_i) f_0(X_j) \kappa(X_i,X_j), \quad && T_2 := \frac{1}{n^2} \sum_{i = 1}^{n} Y_i^2 \kappa(X_i,X_i).
-\end{align*}
-The variances of each statistic can be found by routine computation (see~\cite{ingster2009}), and in particular satisfy the upper bounds
-\begin{align*}
-\Var(T_{1,1}) & \leq \frac{2K}{n^2}, \quad && \Var(T_{1,2}) \overset{\mathrm{(i)}}{\leq} \frac{C}{n}\|f_0\|_P^2\\
-\Var(T_{1,3}) & \overset{\mathrm{(ii)}}{\leq} C\biggl(\frac{K}{n}\|f_0\|_P^4 + \frac{K}{n^2}\|f_0\|_{L^4(\mc{X})}^4\biggr), \quad && \Var(T_{2}) \overset{\mathrm{(iii)}}{\leq} \frac{CK^2}{n^3}\biggl(1 + \|f_0\|_{L^4(\mc{X})}^4\biggr)
-\end{align*}
-where $\mathrm{(i)}-\mathrm{(iii)}$ hold due to local Weyl's law, i.e.~\eqref{eqn:local_weyl}. Upper bounds on Type I and Type II error,
-\begin{equation*}
-\Ebb_0[\wt{\varphi}] \leq \biggl(1 + CK/n^2\biggr)a, \quad \Ebb_{f_0}[1 - \wt{\varphi}] \leq \frac{C(K/n^2 + \|f_0\|_P^2 + K/n \|f_0\|_P^4 + K/n^2 \|f_0\|_{L^4(\mc{X})}^4)}{(\sum_{k = 1}^{K} \dotp{f_0}{\psi_k}_P^2 - \sqrt{2K/an})^2},
-\end{equation*}
-follow from Chebyshev's inequality. It can be verified that so long as
-\begin{equation}
-\label{pf:spectral_series_test}
-\|f_0\|_P^2 \geq C\Biggl(\frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\lambda_{K + 1}^s(\Delta_{P})} + \frac{\sqrt{K}}{n}\biggl(\sqrt{\frac{1}{a}} + \sqrt{\frac{1}{b}}\biggr)\Biggr)
-\end{equation}
-for a sufficiently large constant $C$, then $\Ebb_{f_0}[1 - \wt{\varphi}] \leq b$.
-The two summands in~\eqref{pf:spectral_series_test} are bias and standard deviation terms, respectively.
-When $M^2 \leq n^{-1}$, setting $K = 1$ gives the desired result.
-Otherwise, choosing $K = \floor{M^2n}^{2d/(4s + d)}$ balances these two terms, and leads to~\eqref{eqn:spectral_series_testing}.
-\qed
-% (AG 8/19/21): Same comment as above, regarding the role of M in random design regression testing.
-
-The main takeaway from Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} is that spectral projection methods for nonparametric regression achieve optimal rates of convergence, when the regression function $f_0$ is Sobolev smooth and the design density $p$ is known a priori and satisfies an appropriate notion of smoothness\footnote{The assumption $p \in C^{\infty}(\mc{X})$ could likely be weakened, but since this would not substantially add to the main points of Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, we do not pursue the details further.}.
-As we will see, Laplacian eigenmaps achieves comparable rates of convergence when $p$ is sufficiently smooth but potentially unknown, and in this sense both learns and leverages the design distribution in a way that more classical spectral projection methods do not.
-
-Of course, it is worth pointing out that other methods besides Laplacian eigenmaps are statistically optimal for nonparametric regression even when $p$ is unknown.
-We comment more on some of these in Section~\ref{sec:discussion}, after we have derived our major results regarding Laplacian eigenmaps.
-
-\section{Minimax Optimality of Laplacian Eigenmaps}
+
+The main takeaway from Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing} is that population-level spectral series methods for regression achieve optimal rates of convergence, when the regression function $f_0$ is Sobolev smooth and the design distribution $P$ is known a priori and satisfies an appropriate notion of smoothness.\footnote{The assumption $p \in C^{\infty}(\mc{X})$ could likely be weakened, but since this would not substantially add to the main points of Propositions~\ref{prop:spectral_series_estimation} and~\ref{prop:spectral_series_testing}, we do not pursue the details further.} We reiterate that when the design distribution is unknown, these methods have to be treated as oracle methods, in contrast to PCR-LE.
+As we will see, PCR-LE achieves comparable rates of convergence when $p$ is sufficiently smooth but potentially unknown.
+
+Of course, it is worth pointing out that other methods besides PCR-LE are statistically optimal for nonparametric regression even when $p$ is unknown.
+We comment more on some of these in Section~\ref{sec:discussion}, after we have derived our major results regarding PCR-LE.
+
+\section{Minimax Optimality of PCR-LE}
 \label{sec:minimax_optimal_laplacian_eigenmaps}
 
-In this section we give upper bounds on the error of Laplacian eigenmaps in the flat Euclidean setting, where we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to Model~\ref{def:model_flat_euclidean}.
+In this section we give upper bounds on the error of PCR-LE in the flat Euclidean setting, where we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$ according to Model~\ref{def:model_flat_euclidean}.
 We will divide our theorem statements based on whether the regression function $f_0$ belongs to the first order Sobolev class $H^1(\mc{X})$ or a higher-order Sobolev class ($H_0^{s}(\mc{X})$ for some $s > 1$), since the details of the two settings are somewhat different.
@@
 
-\paragraph{Estimation.} Laplacian eigenmaps depends on the kernel $\eta$ and two tuning parameters, the graph radius $\varepsilon$ and number of eigenvectors $K$.
-We will need to make some assumptions on each.
+\paragraph{Estimation.} PCR-LE depends on the kernel $\eta$ and two tuning parameters, the graph radius $\varepsilon$ and number of eigenvectors $K$. We will need to make some assumptions on each.
 \begin{enumerate}[label=(K\arabic*)]
@@
 \end{enumerate}
-We comment on these assumptions after stating our first main theorem, regarding the estimation error of Laplacian eigenmaps.
+We comment on these assumptions after stating our first main theorem, regarding the estimation error of PCR-LE.
 The proof of this theorem, along with the proofs of all subsequent results, can be found in the Appendix.
@@
 	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H^1(\mc{X},M)$.
-There are constants $c,C$ and $N$ (not depending on $f_0$, $M$ or $n$), such that the following statement holds for all $n \geq N$ and any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_fo}, then
+There are constants $c,C$ and $N$ (not depending on $f_0$, $M$ or $n$), such that the following statement holds for all $n \geq N$ and any $\delta \in (0,1)$: if the PCR-LE estimator $\wh{f}$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_fo}, then
 	\begin{equation}
@@
 	\item \emph{Radius of the Sobolev ball.} 
-	When $M = o(n^{-1/2})$ then computing Laplacian eigenmaps with $K = 1$ achieves the parametric rate $\|\wh{f} - f_0\|_n^2 \lesssim n^{-1}$, and the zero-estimator $\wh{f} = 0$ achieves the better rate $\|\wh{f} - f_0\|_n^2 \lesssim M^2$.
+	When $M = o(n^{-1/2})$ then computing PCR-LE with $K = 1$ achieves the parametric rate $\|\wh{f} - f_0\|_n^2 \lesssim n^{-1}$, and the zero-estimator $\wh{f} = 0$ achieves the better rate $\|\wh{f} - f_0\|_n^2 \lesssim M^2$.
 However, we do not know what the minimax rate is in this regime.
-On the other hand, when $M = \omega(n^{1/d})$, then computing Laplacian eigenmaps with $K = n$ achieves the rate $\|\wh{f} - f_0\|_n^2 \lesssim 1$, which is better than the rate in~\eqref{eqn:spectral_series_estimation}.
+On the other hand, when $M = \omega(n^{1/d})$, then computing PCR-LE with $K = n$ achieves the rate $\|\wh{f} - f_0\|_n^2 \lesssim 1$, which is better than the rate in~\eqref{eqn:spectral_series_estimation}.
 This is because we are evaluating error in-sample rather than out-of-sample.
@@
 
-	\item \emph{In-sample error.} Since the Laplacian eigenmaps estimator is defined only at the design points $\{X_1,\ldots,X_n\}$, we use the empirical norm $\|\cdot\|_n^2$ as our estimation loss.
-We return to the question of out-of-sample estimation later in Section~\ref{sec:out_of_sample}.
-
-	There is one subtlety introduced by the use of in-sample mean squared error.
+	\item \emph{In-sample error.} Since the PCR-LE estimator is defined only at the design points $\{X_1,\ldots,X_n\}$, we use the empirical norm $\|\cdot\|_n^2$ as our estimation loss.
+Depending on the problem at hand, it may be more interesting to consider loss in $L^2(P)$ norm, for instance because this loss is intrinsically tied to the prediction error $\Ebb_{X \sim P}(\wh{f}(X) - f_0(X))^2$.
+In some preliminary analysis, we have considering applying a generic kernel smoother to the PCR-LE estimate, so as to produce a bona-fide function $T\wh{f}: \Rd \to \Reals$ for which the error $\|T\wh{f} - f_0\|_P^2$ is provably on the order of the minimax estimation rate $n^{-2/(2 + d)}$.
+Indeed, our analysis is completely independent of the structure of $\wh{f}$, and could apply to any estimator $\wh{g} \in \Reals^n$ sufficiently close to $f_0$ in empirical norm.
+We also establish that the optimal bandwidth for such a kernel smoother is meaningfully smaller than used for noisy data---which is intuitive, because we are applying it to denoised estimates---so that actually the error  $\|T\wh{f} - f_0\|_P^2$ is dominated by the in-sample mean-squared error $\|\wh{f} - f_0\|_n^2$, plus a negligible extra term.
+We intend to pursue the matter further in future work.
+
+	\item \emph{Meaning of pointwise evaluation.} There is one subtlety introduced by the use of in-sample mean squared error.
 Since elements $f \in H^s(\mc{X})$ are equivalence classes, defined only up to a set of measure zero, one cannot really speak of the pointwise evaluation $f_0(X_i)$, as we do by defining our target of estimation to be $(f_0(X_1),\ldots,f_0(X_n))$, until one selects a representative of each equivalence class $f$.
@@
 The lower bound on $\varepsilon$ imposed by~\eqref{eqn:radius_fo} is on the order of the connectivity threshold, the smallest radius for which the resulting graph will still be connected with high probability.
-On the other hand, as we will see in Section~\ref{subsec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K$ is of at least the same order as the continuum eigenvalue $\lambda_K(\Delta_P)$; this is essential in order to obtain a tight upper bound on the bias of $\wh{f}$.
+On the other hand, as we will see in Section~\ref{subsec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K$ is of at least the same order as the continuum eigenvalue $\rho_K$; this is essential in order to obtain a tight upper bound on the bias of $\wh{f}$.
 Finally, we set $K = \floor{(M^2n)^{d/(2 + d)}}$ (when possible) to optimally trade-off bias and variance.
 
-	In practice, one typically tunes hyper-parameters by cross-validation.
-However, because the estimator $\wh{f}$ is defined only in-sample, neither cross-validation nor any other sample-splitting technique can be used to tune parameters for Laplacian eigenmaps.
+	% Eliminated the following chunk because I'm no longer sure what to say.
+	% In practice, one typically tunes hyper-parameters by cross-validation.
+However, because the estimator $\wh{f}$ is defined only in-sample, neither cross-validation nor any other sample-splitting technique can be used to tune parameters for PCR-LE.
 We return to this issue in Section~\ref{sec:out_of_sample}, when we propose an out-of-sample extension of $\wh{f}$.
@@
 	% (AG 8/27/21): Ryan, you asked about AIC, but I'm not clear about whether you want me to (a) make some remark about it, (b) actually go through the work of giving guarantees, or (c) were just curious.
+
 	\item \emph{High-probability guarantees}.
@@
 Suppose additionally $f_0 \in H^1(\mc{X};M)$, and that $d < 4$.
-Then there exist constants $C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n \geq N$: if the Laplacian eigenmaps test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_fo}, and if $f_0$ satisfies
+Then there exist constants $C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n \geq N$: if the PCR-LE test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_fo}, and if $f_0$ satisfies
 	\begin{equation}
@@
 We now consider the situation where the regression function displays some higher-order regularity, $f_0 \in H_0^s(\mc{X};M)$.
-We show that Laplacian eigenmaps methods continue to be optimal for all orders of $s$, as long as the design density is itself also sufficiently regular, $p \in C^{s - 1}(\mc{X})$.
+We show that the PCR-LE estimator and test continue to be optimal for all orders of $s$, as long as the design density is itself also sufficiently regular, $p \in C^{s - 1}(\mc{X})$.
 In estimation, this is the case for any dimension $d$, whereas in testing it is the case only when $d \leq 4$.
@@
 	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H_0^s(\mc{X},M)$ and $p \in C^{s - 1}(\mc{X})$.
-There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_ho}, then
+There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the PCR-LE estimator $\wh{f}$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_ho}, then
 	\begin{equation}
@@
 \end{theorem}
-Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that in the flat Euclidean setting Laplacian eigenmaps is a minimax rate-optimal estimator over Sobolev classes, for all values of $s$ and $d$.
+Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, in combination with Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, implies that in the flat Euclidean setting PCR-LE is a minimax rate-optimal estimator over Sobolev classes, for all values of $s$ and $d$.
 Some other remarks:
@@
 Likewise, certain estimators are ``off the table'', most notably RKHS-based methods such as thin-plate splines of degree $k \leq d/2$.
-Nevertheless, for random design regression with error measured in squared $L^2(P)$-norm, the spectral projection estimator $\wt{f}$ obtains the standard minimax rates $n^{-2s/(2s + d)}$ for all values of $s$ and $d$.
-Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and \ref{thm:laplacian_eigenmaps_estimation_ho} show that the same is true with respect to Laplacian eigenmaps, when error is measured in squared $L^2(P_n)$-norm.
+Nevertheless, for random design regression with error measured in squared $L^2(P)$-norm, the population-level spectral series estimator $\wt{f}$ obtains the standard minimax rates $n^{-2s/(2s + d)}$ for all values of $s$ and $d$.
+Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and \ref{thm:laplacian_eigenmaps_estimation_ho} show that the same is true with respect to PCR-LE, when error is measured in squared $L^2(P_n)$-norm.
 	\item \emph{Smoothness of design density}.
-As promised,  Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} shows that Laplacian eigenmaps achieves optimal rates of convergence so long as the unknown design density $p$ is sufficiently smooth, $p \in C^{s - 1}(\mc{X})$.
+As promised,  Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} shows that PCR-LE achieves optimal rates of convergence so long as the unknown design density $p$ is sufficiently smooth, $p \in C^{s - 1}(\mc{X})$.
 The requirement $p \in C^{s - 1}(\mc{X})$ is essential to showing that $\wh{f}$ enjoys the faster minimax rates of convergence when $s > 1$,  as we discuss in Section~\ref{subsec:analysis}.
@@
 To address this issue, in Appendix~\ref{sec:computational_considerations} we review some approaches to \emph{spectral sparsification}, in which one efficiently computes a sparse graph $\wc{G}$ that approximates $G$ in a spectral sense.
-The hope is that the Laplacian eigenmaps estimator $\wc{f}$ computed with respect to the sparsified graph $\wc{G}$ has similar statistical properties as $\wh{f}$, while being much faster to compute.
+The hope is that the PCR-LE estimator $\wc{f}$ computed with respect to the sparsified graph $\wc{G}$ has similar statistical properties as $\wh{f}$, while being much faster to compute.
 To that end, we provide upper bounds on $\|\wc{f} - f_0\|_n^2$, which show that under mild conditions on $\wc{G}$---provably achieved by many spectral sparsification algorithms---the estimator $\wc{f}$ achieves the same rates of convergence as $\wh{f}$.
@@
 Suppose additionally $f_0 \in H_0^s(\mc{X},M)$, that $p \in C^{s-1}(\mc{X})$, and that $d \leq 4$.
-Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n \geq N$: if the Laplacian eigenmaps test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_ho}, and if $f_0$ satisfies
+Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n \geq N$: if the PCR-LE test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_ho}, and if $f_0$ satisfies
 	\begin{equation}
@@
 	Under the same setup as Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, but with $4 < d < 4s$.
-If the Laplacian eigenmaps test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, number of eigenvectors $K$ satisfying~\eqref{eqn:eigenvector_testing_ho}, and $\varepsilon = (M^2n)^{-1/(2(s - 1) + d)}$, and if
+If the PCR-LE test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_flat_euclidean}, number of eigenvectors $K$ satisfying~\eqref{eqn:eigenvector_testing_ho}, and $\varepsilon = (M^2n)^{-1/(2(s - 1) + d)}$, and if
 	\begin{equation}
@@
 \end{theorem}
-Focusing on the special case where $M \asymp 1$, Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} says that $\varphi$ has have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2(s - 1) + d)}$ and $4 < d < 4s$.
+Focusing on the special case where $M \asymp 1$, Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} says that $\varphi$ has small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2(s - 1) + d)}$ and $4 < d < 4s$.
 This is smaller than the estimation rate $n^{-2s/(2s + d)}$, but larger than the minimax squared critical radius $n^{-4s/(4s + d)}$.
 
-% (AG 8/30/21): Is the below chunk helpful? Or not intuitive enough to be reasonable.
-It is intuitively reasonable that data-dependent spectral projection methods, such as Laplacian eigenmaps, should have more difficulty achieving the minimax rates of convergence for testing, as opposed to estimation.
-Optimally balancing the bias and variance of spectral projection tests (for instance~\eqref{pf:spectral_series_test}) requires choosing more and higher-frequency eigenvectors than in estimation.
-It is known~\citep{trillos2019} that higher-frequency graph Laplacian eigenvectors---which correspond to larger eigenvalues---are meaningfully worse approximations of their population-level limits.
+% (AG 8/30/21): Is the below chunk helpful? Or not intuitive enough.
+It is intuitively reasonable that data-dependent spectral series methods, such as PCR-LE, should have more difficulty achieving the minimax rates of convergence for testing, as opposed to estimation.
+Optimally balancing the bias and variance of the PCR-LE test requires choosing more and higher-frequency eigenvectors than in estimation.
+Recall from~\eqref{eqn:spectral_gap} that higher-frequency graph Laplacian eigenvectors---which correspond to larger eigenvalues---are meaningfully worse approximations of their population-level limits.
 Put more simply, minimax optimal testing when $d > 4$ appears to require stronger approximation properties than graph Laplacian eigenvectors can satisfy.
@@
 We now outline the high-level strategy we follow when proving each of Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}.
-We analyze the estimation error of $\wh{f}$, and the testing error of $\varphi$, by first conditioning on the design points $\{X_1,\ldots,X_n\}$ and deriving \emph{design-dependent} bias and variance terms.
-For estimation, we have that with probability at least $1 - \exp(-K)$,
+We analyze the estimation error of $\wh{f}$, and the testing error of $\varphi$, by first conditioning on the design points $\{X_1,\ldots,X_n\}$ and deriving \emph{design-dependent} bias and variance terms, as outlined in~\eqref{eqn:bias_variance_estimation} and~\eqref{eqn:bias_estimation}.
+For estimation, we show that with probability at least $1 - \exp(-K)$,
 \begin{equation}
@@
 \end{equation}
-For testing, we have that $\varphi$ (which is a level-$a$ test by construction) also has small Type II Error, $\Ebb_{f_0}[1 - \phi] \leq b/2$, if 
+For testing, we show that $\varphi$ (which is a level-$a$ test by construction) also has small Type II Error, $\Ebb_{f_0}[1 - \varphi] \leq b/2$, if 
 \begin{equation}
@@
 \end{equation}
-These design-dependent bias-variance decompositions are reminiscent of the more classical bias-variance decompositions typical in the analysis of population-level spectral projection methods, but different in certain key respects.
-Comparing~\eqref{eqn:estimation_biasvariance} and~\eqref{eqn:testing_biasvariance} to~\eqref{pf:spectral_series_estimation_3} and~\eqref{pf:spectral_series_test}, we see that two continuum objects in the latter pair of bounds, the Sobolev norm $\|f_0\|_{\mc{H}^s(\mc{X})}^2$ and the eigenvalue $\lambda_k(\Delta_P)$, have been replaced by graph-based analogues: the graph Sobolev seminorm $\dotp{L_{n,\varepsilon}^sf_0}{f_0}_n$ and the graph Laplacian eigenvalue $\lambda_k^s$.
+These design-dependent bias-variance decompositions are reminiscent of the more classical bias-variance decompositions typical in the analysis of population-level spectral series methods (for instance\eqref{pf:spectral_series_estimation_3} and~\eqref{pf:spectral_series_test}), but different in certain key respects.
+Comparing~\eqref{eqn:estimation_biasvariance} and~\eqref{eqn:testing_biasvariance} to~\eqref{pf:spectral_series_estimation_3} and~\eqref{pf:spectral_series_test}, we see that two continuum objects in the latter pair of bounds, the Sobolev norm $\|f_0\|_{\mc{H}^s(\mc{X})}^2$ and the eigenvalue $\rho_k^s$, have been replaced by graph-based analogues: the graph Sobolev seminorm $\dotp{L_{n,\varepsilon}^sf_0}{f_0}_n$ and the graph Laplacian eigenvalue $\lambda_k^s$.
 These latter quantities, along with the empirical squared norm $\|f_0\|_n^2$, are random variables that depend on the random design points $\{X_1,\ldots,X_n\}$.
@@
 \paragraph{Estimates of graph Sobolev seminorms.}
-In Proposition~\ref{prop:graph_seminorm_fo} we restate an upper bound on the Dirichlet energy $\dotp{L_{n,\varepsilon}f}{f}_n$ from~\cite{green2021}. 
+In Proposition~\ref{prop:graph_seminorm_fo} we restate an upper bound on the first-order graph Sobolev semi-norm $\dotp{L_{n,\varepsilon}f}{f}_n$ from~\cite{green2021}. 
 \begin{proposition}[Lemma 1 of~\cite{green2021}]
@@
 	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f \in H^1(\mc{X})$.
-There exist constants $c,C$ that do not depend on $f$ or $n$ such that the following statement holds for any $\delta \in (0,1)$: if $\eta$ satisfies~\ref{asmp:kernel} and $\varepsilon < c$, then
+There exist constants $c,C$ that do not depend on $f$ or $n$ such that the following statement holds for any $\delta \in (0,1)$: if $\eta$ satisfies~\ref{asmp:kernel_flat_euclidean} and $\varepsilon < c$, then
 	\begin{equation}
@@
 \begin{enumerate}
-	\item First we note that $\dotp{L_{n,\varepsilon}^s f}{f}_n$ is itself a biased estimate of the non-local seminorm $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$.
-Specifically, $\dotp{L_{n,\varepsilon}^s f}{f}_n$ is a $V$-statistic, meaning it is the sum of an unbiased estimator of $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$ (in other words, a $U$-statistic) plus some higher-order, pure bias terms.
+	\item \emph{Bound on pure bias terms.} First we note that $\dotp{L_{n,\varepsilon}^s f}{f}_n$ is itself a biased estimate of the non-local seminorm $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$.
+This is because $\dotp{L_{n,\varepsilon}^s f}{f}_n$ is a $V$-statistic, meaning it is the sum of an unbiased estimator of $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$ (in other words, a $U$-statistic) plus some higher-order, pure bias terms.
 We show that these pure bias terms are negligible when $\varepsilon = \omega(n^{-1/(2(s - 1) + d)})$.
-	\item For $x$ sufficiently far from the boundary of $\mc{X}$---precisely $x \in \mc{X}$ such that $B(x,j\varepsilon) \subseteq \mc{X}$---we show that $L_{P,\varepsilon}^jf(x) \to \sigma_{\eta}^j \Delta_P^jf(x)$ as $\varepsilon \to 0$.
+	\item \emph{Convergence in the interior.} For $x$ sufficiently far from the boundary of $\mc{X}$---precisely $x \in \mc{X}$ such that $B(x,j\varepsilon) \subseteq \mc{X}$---we show that $L_{P,\varepsilon}^jf(x) \to \sigma_{\eta}^j \Delta_P^jf(x)$ as $\varepsilon \to 0$.
 Here $j = (s - 1)/2$ when $s$ is odd and $j = (s - 2)/2$ when $s$ is even.
 This step bears some resemblance to the analysis of the bias term in kernel smoothing, and requires that $p \in C^{s-1}(\mc{X})$.
-	\item On the other hand for $x$ sufficiently near the boundary of $\mc{X}$, $L_{P,\varepsilon}^jf(x)$ does not in general converge to $\sigma_{\eta}^j\Delta_P^jf(x)$.
+	\item \emph{Boundedness at the boundary.} On the other hand for $x$ sufficiently near the boundary of $\mc{X}$, $L_{P,\varepsilon}^jf(x)$ does not in general converge to $\sigma_{\eta}^j\Delta_P^jf(x)$.
 Instead, we use the zero-trace property of $f$ to show that $L_{P,\varepsilon}^jf(x)$ is small.
-	\item Finally, we combine the results of the previous two steps to deduce an upper bound on $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$ in terms of the squared Sobolev norm $\|f\|_{H^s(\mc{X})}^2$.
-Roughly speaking, when $s$ is odd, $\dotp{L_{P,\varepsilon}^sf}{f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}^jf) \approx \sigma_{\eta}^{2j}E_{P,\varepsilon}(\Delta_P^jf)$, whereas when $s$ is even $\dotp{L_{P,\varepsilon}^sf}{f}_P = \|L_{P,\varepsilon}L_{P,\varepsilon}^{j}f\|_{P}^2 \approx \sigma_{\eta}^{2j}\|L_{P,\varepsilon} \Delta_Pf\|_P^2$.
-Reasoning in this way, we can translate estimates of $L_{P,\varepsilon}^jf$ into an upper bound on the order-$s$ non-local Sobolev seminorm, even though $s > j$.
+	\item \emph{Putting together the pieces.} Finally, we combine the results of the previous two steps to deduce an upper bound on $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$ in terms of the squared Sobolev norm $\|f\|_{H^s(\mc{X})}^2$.
+The nature of this last step depends on whether $s$ is an even or an odd integer.
+Roughly speaking, when $s$ is odd, letting $j = (s - 1)/2$, we show that
+	\begin{equation*}\dotp{L_{P,\varepsilon}^sf}{f}_P = \dotp{L_{P,\varepsilon}L_{P,\varepsilon}^jf}{L_{P,\varepsilon}^jf}_P \approx \sigma_{\eta}^{2j}\dotp{L_{P,\varepsilon}\Delta_P^jf}{\Delta_P^j}_{P} \lesssim \sigma_{\eta}^{2j + 1} \dotp{\Delta_P^jf}{f}_{P}.
+	\end{equation*}
+	When $s$ is even, letting $j = (s - 2)/2$, we show that
+	\begin{equation*}
+	\dotp{L_{P,\varepsilon}^sf}{f}_P = \|L_{P,\varepsilon}L_{P,\varepsilon}^{j}f\|_{P}^2 \approx \sigma_{\eta}^{2j}\|L_{P,\varepsilon} \Delta_P^jf\|_P^2 \lesssim \|\Delta_P^{j + 1}f\|_P^2.
+	\end{equation*}
+	In either case, the desired upper bound~\eqref{eqn:graph_seminorm_ho} follows from the boundedness of the density $p$.
 \end{enumerate}
-Together, these steps establish Proposition~\ref{prop:graph_seminorm_ho}.
-It is worth pointing out that we do not try to establish the pointwise estimate $L_{P,\varepsilon}^sf \to \sigma_{\eta}^s\Delta_{P}^sf$ in $L^2(P)$.
+It is worth pointing out that we do not need to establish the pointwise estimate $L_{P,\varepsilon}^sf \to \sigma_{\eta}^s\Delta_{P}^sf$ in $L^2(P)$.
 If we had such an estimate, it would immediately follow that $\dotp{L_{P,\varepsilon}^sf}{f}_{P} \to \sigma_{\eta}^s\dotp{\Delta_P^sf}{f}_{P}$.
@@
 \paragraph{Neighborhood graph eigenvalues.}
-On the other hand, several recent works \citep{burago2014,garciatrillos18,calder2019} have analyzed the convergence of $\lambda_{k}$ towards $\lambda_{k}(\Delta_P)$.
-They provide explicit bounds on the relative error $|\lambda_{k} - \lambda_{k}(\Delta_P)|/\lambda_{k}(\Delta_P)$, which show that the relative error is small for sufficiently large $n$ and small $\varepsilon$.
-Crucially, these guarantees hold simultaneously for all $1 \leq k \leq K$ as long as $\lambda_{K}(\Delta_P) = O(\varepsilon^{-2})$.
+On the other hand, several recent works \citep{burago2014,garciatrillos18,calder2019} have analyzed the convergence of $\lambda_{k}$ towards $\rho_k$.
+They provide explicit bounds on the relative error $|\lambda_{k} - \rho_k|/\rho_k$, which show that the relative error is small for sufficiently large $n$ and small $\varepsilon$.
+Crucially, these guarantees hold simultaneously for all $1 \leq k \leq K$ as long as $\\rho_K = O(\varepsilon^{-2})$.
 These results are actually stronger than are necessary to establish Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho}---in order to get rate-optimality, we need only show that for the relevant values of $K$, $\lambda_{K}/\lambda_K(P) = \Omega_P(1)$---but unfortunately they all assume $P$ is supported on a manifold without boundary (i.e.
@@
 In the case where $\mc{X}$ is assumed to have a boundary, the graph Laplacian $L_{n,\varepsilon}$ is a reasonable approximation of the operator $\Delta_P$ only at points $x \in \mc{X}$ for which $B(x,\varepsilon) \subseteq \mc{X}$.
-In contrast, at points $x$ near the boundary of $\mc{X}$, the graph Laplacian is known to approximate a different operator altogether \citep{belkin2012}.\footnote{This is directly related to the boundary bias of kernel smoothing, since the graph Laplacian can be viewed as a kernel-based estimator of $\Delta_P$.} This renders analysis of $\lambda_k$ substantially more challenging, since its continuum limit is not $\lambda_k(\Delta_P)$.
-Rather than analyzing the convergence of $\lambda_k$, we will instead use Lemma~2 of \cite{green2021}, whose assumptions match our own, and who give a weaker bound on the ratio $\lambda_k/\lambda_k(\Delta_P)$ that will nevertheless suffice for our purposes.
+In contrast, at points $x$ near the boundary of $\mc{X}$, the graph Laplacian is known to approximate a different operator altogether \citep{belkin2012}.\footnote{This is directly related to the boundary bias of kernel smoothing, since the graph Laplacian can be viewed as a kernel-based estimator of $\Delta_P$.} This renders analysis of $\lambda_k$ substantially more challenging, since its continuum limit is not $\rho_k$.
+Rather than analyzing the convergence of $\lambda_k$, we will instead use Lemma~2 of \cite{green2021}, whose assumptions match our own, and who give a weaker bound on the ratio $\lambda_k/\rho_k$ that will nevertheless suffice for our purposes.
 
@@
 	\label{eqn:graph_eigenvalue}
-	\lambda_k \geq c \cdot \min\Bigl\{\lambda_k(\Delta_P), \frac{1}{\varepsilon^{2}} \Bigr\} \quad \textrm{for all $1 \leq k \leq n$,}
+	\lambda_k \geq c \cdot \min\Bigl\{\rho_k, \frac{1}{\varepsilon^{2}} \Bigr\} \quad \textrm{for all $1 \leq k \leq n$,}
 	\end{equation}
@@
 By our assumptions on $P$, $\lambda_0(\Delta_P) = \lambda_0 = 0$.
-Furthermore, Weyl's Law~\eqref{eqn:weyl} tells us that under Model~\ref{def:model_flat_euclidean}, $k^{2/d} \lesssim \lambda_{k}(\Delta_P) \lesssim k^{2/d}$ for all $k \in \mathbb{N}, k > 1$.
+Furthermore, Weyl's Law~\eqref{eqn:weyl} tells us that under Model~\ref{def:model_flat_euclidean}, $k^{2/d} \lesssim \rho_k \lesssim k^{2/d}$ for all $k \in \mathbb{N}, k > 1$.
 Combining these statements with~\eqref{eqn:graph_eigenvalue}, we conclude that $\lambda_{K} = \Omega_P(K^{2/d})$ so long as $K \lesssim \varepsilon^{-d}$.
@@
 % (AG 8/29/21): Following chunk commented out but not deleted, in case it proves useful at a later point.
-% We conclude this section by pointing out that, at first glance, it may be somewhat remarkable that Laplacian eigenmaps can ever attain optimal rates of convergence.
+% We conclude this section by pointing out that, at first glance, it may be somewhat remarkable that PCR-LE can ever attain optimal rates of convergence.
 The sharpest known results \citep{calder2019,cheng2021} show that the squared error with which a graph Laplacian eigenvector approximates its continuum limit, $\|v_k - \psi_k\|_n^2$,  converges to $0$ at a rate of $C_k n^{-2/(4 + d)}$, for a constant $C_k \to \infty$ as $k \to \infty$.
-For any fixed $k$, this is slower than the optimal estimation or testing rates, and when the number of eigenvectors $K$ increases with $n$, as is necessary to optimally balance bias and variance of Laplacian eigenmaps, the issue only gets worse.
+For any fixed $k$, this is slower than the optimal estimation or testing rates, and when the number of eigenvectors $K$ increases with $n$, as is necessary to optimally balance bias and variance of PCR-LE, the issue only gets worse.
 Moreover, there are reasons to believe that this upper bound on $\|v_k - \psi_k\|_n^2$ is unimprovable (see~\cite{cheng2021}).
-If so, this suggests that Laplacian eigenmaps can learn regression function more efficiently, from a statistical perspective, than Laplacian eigenvectors can learn population-level eigenfunctions.
+If so, this suggests that PCR-LE can learn regression function more efficiently, from a statistical perspective, than Laplacian eigenvectors can learn population-level eigenfunctions.
 Put more simply: regression is a forgiving problem.
@@
 In this setting, it is known that the minimax rates depend only on the intrinsic dimension $m$; more specifically,~\cite{bickel2007,ariascastro2018} show that for functions with H\"{o}lder smoothness $s$, the minimax estimation rate is $n^{-2s/(2s + m)}$ and the testing rate is $n^{-4s/(4s + m)}$.\footnote{Although~\cite{ariascastro2018} considers density testing, usual arguments regarding equivalence of experiments~\citep{brown96} imply that the same rates apply to regression testing.} On the other hand, a theory has been developed \citep{niyogi2008finding,belkin03,belkin05,niyogi2013,balakrishnan2012minimax,balakrishnan2013cluster} establishing that the neighborhood graph $G$ can ``learn'' the manifold $\mc{X}$ in various senses, so long as $\mc{X}$ is locally linear.
-We build on this work by showing that when $P$ is supported on a manifold $\mc{X}$ and $f_0 \in H^s(\mc{X})$, Laplacian eigenmaps achieves the sharper minimax estimation and testing rates.
+We build on this work by showing that when $P$ is supported on a manifold $\mc{X}$ and $f_0 \in H^s(\mc{X})$, PCR-LE achieves the sharper minimax estimation and testing rates.
 
@@
 Unlike in the flat-Euclidean case, since Model~\ref{def:model_manifold} assumes that $\mc{X}$ is without boundary it is easy to deal with the first-order $(s = 1)$ and higher-order $(s > 1)$ cases all at once.
-A more important distinction between the results of this section and those of Section~\ref{sec:minimax_optimal_laplacian_eigenmaps} is that we will establish Laplacian eigenmaps is optimal only when the regression function $f_0 \in H^s(\mc{X};M)$ for $s \in \{1,2,3\}$.
+A more important distinction between the results of this section and those of Section~\ref{sec:minimax_optimal_laplacian_eigenmaps} is that we will establish PCR-LE is optimal only when the regression function $f_0 \in H^s(\mc{X};M)$ for $s \in \{1,2,3\}$.
 Otherwise, this section will proceed in a similar fashion to Section~\ref{sec:higher_order_sobolev_classes}.
@@
 	Suppose Model~\ref{def:model_manifold}, and additionally $f_0 \in H^s(\mc{X},M)$ and $p \in C^{s - 1}(\mc{X})$ for $s \leq 3$.
-There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_manifold}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_manifold}, then
+There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all for all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the PCR-LE estimator $\wh{f}$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_manifold}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_estimation_manifold}, then
 	\begin{equation}
@@
 Suppose additionally $f_0 \in H^s(\mc{X},M)$, that $p \in C^{s-1}(\mc{X})$, and that $s \leq 3$ and $m \leq 4$.
-Then there exist constants $c$, $C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_manifold}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_manifold}, and if $f_0$ satisfies
+Then there exist constants $c$, $C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the PCR-LE test $\varphi$ is computed with a kernel $\eta$ satisfying~\ref{asmp:kernel_manifold}, and parameters $\varepsilon$ and $K$ satisfying~\ref{asmp:parameters_testing_manifold}, and if $f_0$ satisfies
 	\begin{equation}
@@
 \end{theorem}
-Focusing on the case $M \asymp 1$,\footnote{To the best of our knowledge, the minimax rates for general $M$ in the manifold setting have not been worked out.} the upper bounds in Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold} imply that Laplacian eigenmaps attain the optimal rates of convergence over Sobolev balls $H^s(\mc{X})$ for $s \in \{1,2,3\}$.
-
-Unlike in the full-dimensional case, in the manifold setting our upper bounds on the estimation and testing error of Laplacian eigenmaps do not match the minimax rate when $s \geq 4$.
-In this case, the containment $H^s(\mc{X};1) \subset H^{3}(\mc{X};1)$ implies that the Laplacian eigenmaps estimator $\wh{f}$ has in-sample mean-squared error of at most on the order of $n^{-6/(6 + m)}$, and that the Laplacian eigenmaps test has small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-12/(12 + m)}$; however, these are slower than the minimax rates.
+Focusing on the case $M \asymp 1$,\footnote{To the best of our knowledge, the minimax rates for general $M$ in the manifold setting have not been worked out.} the upper bounds in Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold} imply that PCR-LE attain the optimal rates of convergence over Sobolev balls $H^s(\mc{X})$ for $s \in \{1,2,3\}$.
+
+Unlike in the full-dimensional case, in the manifold setting our upper bounds on the estimation and testing error of PCR-LE do not match the minimax rate when $s \geq 4$.
+In this case, the containment $H^s(\mc{X};1) \subset H^{3}(\mc{X};1)$ implies that the PCR-LE estimator $\wh{f}$ has in-sample mean-squared error of at most on the order of $n^{-6/(6 + m)}$, and that the PCR-LE test has small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-12/(12 + m)}$; however, these are slower than the minimax rates.
 
@@
 However this is not an option for us, as we assume $\mc{X}$---and in particular its geodesics---are unknown.
-Likewise, a spectral projection estimator using eigenfunctions of the manifold Laplace-Beltrami operator, will achieve the minimax rate for all values of $s$ and $m$; but this is undesirable for the same reason---we do not want to assume that $\mc{X}$ is known.
-It is not clear whether this gap between spectral projection and Laplacian eigenmaps estimators is real, or a product of loose upper bounds.
+Likewise, a population-level spectral series estimator using eigenfunctions of the manifold Laplace-Beltrami operator, will achieve the minimax rate for all values of $s$ and $m$; but this is undesirable for the same reason---we do not want to assume that $\mc{X}$ is known.
+It is not clear whether this gap between population-level spectral series regression and the PCR-LE estimator is real, or a product of loose upper bounds.
 
 Finally, as in the full-dimensional case, when the intrinsic dimension $m > 4$ we cannot choose the graph radius $\varepsilon$ and number of eigenvectors $K$ to optimally balance bias and variance.
-Instead, reasoning as in the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} shows that when $1 \leq s \leq 3$, the Laplacian eigenmaps test has critical radius as given by~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_ho_suboptimal}, but with the ambient dimension $d$ replaced by $m$.
+Instead, reasoning as in the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} shows that when $1 \leq s \leq 3$, the PCR-LE test has critical radius as given by~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_ho_suboptimal}, but with the ambient dimension $d$ replaced by $m$.
 % (SB via AG): Maybe some comment on the difference between testing and estimation.
@@
 	\label{eqn:graph_eigenvalue_manifold}
-	\lambda_k \geq c \cdot \min\Bigl\{\lambda_k(\Delta_P), \frac{1}{\varepsilon^{2}} \Bigr\} \quad \textrm{for all $1 \leq k \leq n$,}
+	\lambda_k \geq c \cdot \min\Bigl\{\rho_k, \frac{1}{\varepsilon^{2}} \Bigr\} \quad \textrm{for all $1 \leq k \leq n$,}
 	\end{equation}
@@
 
-\section{Out-of-sample error}
-\label{sec:out_of_sample}
-Sections~\ref{sec:minimax_optimal_laplacian_eigenmaps} and~\ref{sec:manifold_adaptivity} show that $\wh{f}$ is a minimax optimal estimator over Sobolev spaces, when the error is measured \emph{in-sample}---that is, in $L^2(P_n)$ norm.
-In this section we consider \emph{out-of-sample} error---measured in $L^2(P)$ norm---which is the more typical metric used in the study of nonparametric regression with random design.
-
-Of course, the Laplacian eigenmaps estimator is only defined at the design points $\{X_1,\ldots,X_n\}$, and to measure error in $L^2(P)$ norm we must first extend it to the entire domain $\mc{X}$.
-We propose a simple method, kernel smoothing, to do the job.
-The method can applied to any estimator defined at the design points, including Laplacian eigenmaps, and we show that a smoothed version of our original estimator $\wh{f}$ has optimal $L^2(P)$ error.
-For simplicity, in this section we will stick to the flat Euclidean setting, where $(X_1,Y_1),\ldots,(X_n,Y_n)$ are observed according to Model~\ref{def:model_flat_euclidean}.
-
-\paragraph{Extension by kernel smoothing.}
-We now formally define our approach to extension by kernel smoothing.
-For a kernel function $\psi(\cdot): [0,\infty) \to (-\infty,+\infty)$, bandwidth $h > 0$, and a distribution $Q$, the \emph{Nadaraya-Watson kernel smoother} $T_{Q,h}$ is given by
-\begin{equation*}
-\bigl(T_{Q,h}f)(x) := 
-\begin{dcases*}
-\frac{1}{d_{Q,h}(x)} \int_{\Omega} f(z)\psi\biggl(\frac{\|z - x\|}{h}\biggr) \,dQ(z), & \textrm{if $d_{Q,h}(x) > 0$,} \\
-0, &\textrm{otherwise,}
-\end{dcases*}
-\end{equation*}
-where $d_{Q,h}(x) := \int_{\Omega} \psi\bigl(\|z - x\|/h\bigr) \,dQ(z)$.
-For convenience, we will write $T_{n,h}f(x) := T_{P_n,h}f(x)$, and $d_{n,h}(x) := n \cdot d_{P_n,h}(x)$.
-We extend the Laplacian eigenmaps estimator by passing the kernel smoother $T_{n,h}$ over it, that is we consider the estimator $T_{n,h}\wh{f}$, which is defined at every $x \in \mc{X}$ (indeed, at every $x \in \Rd$).
-Note that ``extension'' here is a slight abuse of nomenclature, since $T_{n,h}\wh{f}$ and $\wh{f}$ may not agree in-sample.
-
-\paragraph{Out-of-sample error of kernel smoothed Laplacian eigenmaps.}
-
-In Lemma~\ref{lem:kernel_smoothing_insample}, we consider an arbitrary estimator $\wc{f} \in \Reals^n$.
-We show that the out-of-sample error $\|T_{n,h}\wc{f} - f_0\|_P^2$ can be upper bounded by three terms--- (a constant times) the in-sample error $\|\wc{f} - f_0\|_n^2$, and variance and bias terms that arise naturally in the analysis of kernel smoothing over noiseless data.
-We shall assume the following conditions on $\psi$ and $h$.
-\begin{enumerate}[label=(K\arabic*)]
-	\setcounter{enumi}{2}
-	\item
-	\label{asmp:kernel}
-	The kernel function $\psi$ is supported on a subset of $[0,1]$. Additionally, $\psi$ is Lipschitz continuous on $[0,1]$, and is normalized so that
-	\begin{equation*}
-	\int_{-\infty}^{\infty} \psi(|z|) \,dz = 1.
-	\end{equation*}
-\end{enumerate}
-\begin{enumerate}[label=(P\arabic*)]
-	\setcounter{enumi}{6}
-	\item
-	\label{asmp:bandwidth}
-	For constants $c_0$ and $C_0$, the bandwidth parameter $h$ satisfies
-	\begin{equation*}
-	C_0\biggl(\frac{\log(1/h)}{n}\biggr)^{1/d} \leq h \leq c_0.
-	\end{equation*}
-\end{enumerate}
-\begin{lemma}
-	\label{lem:kernel_smoothing_insample}
-	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f_0 \in H^1(\mc{X};M)$ and $p \in C^1(\mc{X})$.
-If the kernel smoothing estimator $T_{n,h}\wc{f}$ is computed with a kernel $\psi$ satisfying~\ref{asmp:kernel} and bandwidth $h$ satisfying~\ref{asmp:bandwidth}, it holds that
-	\begin{equation}
-	\label{eqn:kernel_smoothing_insample}
-	\|T_{n,h}\wc{f} - f_0\|_P^2 \leq C\biggl(\|\wc{f} - f_0\|_n^2 + \frac{1}{\delta} \cdot \frac{h^2}{nh^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{P,h}f_0 - f_0\|_P^2\biggr),
-	\end{equation}
-	with probability at least $1 - \delta - Ch^d\exp\{-Cnh^d\}$. 
-\end{lemma}
-Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data, by a factor of $h^2$.
-On the other hand the bias term is typical, and when $\psi$ is an order-$s$ kernel a standard analysis (see Lemma~\ref{lem:kernel_smoothing_bias}) shows that $\|T_{P,h}f_0 - f_0\|_P^2 \lesssim |f|_{H^s(\mc{X})}^2 h^{2s}$.
-\begin{enumerate}[label=(K\arabic*)]
-	\setcounter{enumi}{3}
-	\item
-	\label{asmp:ho_kernel}
-	The kernel function $\psi$ is an order-$s$ kernel, meaning it satisfies $\int_{-\infty}^{\infty} \psi(|z|) \,dz = 1$, and if $s \geq 2$,
-	\begin{equation*}
-	\int_{-\infty}^{\infty} z^j \psi(|z|) \,dz = 0 ~~\textrm{for}~j = 1,\ldots, s + d - 2, \quad \textrm{and}~ \int_{-\infty}^{\infty} z^{s + d - 1} \psi(|z|) \,dz < \infty. 
-	\end{equation*}
-\end{enumerate}
-Kernels satisfying~\ref{asmp:ho_kernel} can be constructed using linear combinations of (truncated) polynomials in a standard way; for instance, the boxcar kernel $\psi(z) = (1/2)\cdot\1\{|z| \leq 1\}$ is an order-$2$ kernel, and $\psi(z) = (9/26 - 15z^2/52)\cdot\1\{|z| \leq 1\}$ is an order-$3$ kernel.
-
-Choosing $h \asymp n^{-1/(2(s - 1) + d)}$ balances the kernel smoothing bias and variance terms in~\eqref{eqn:kernel_smoothing_insample}, and implies that for $f_0 \in H^s(\mc{X};M)$,
-\begin{equation}
-\label{eqn:kernel_smoothing_insample2}
-\|T_{n,h}\wc{f} - f_0\|_P^2 \leq C\biggl(\|\wc{f} - f_0\|_n^2 + \frac{M^2}{\delta}n^{-2s/(2(s - 1) + d)}\biggr).
-\end{equation}
-This analysis tells us that the additional error incurred by passing a kernel smoother over an in-sample estimator $\wc{f}$ is negligible compared to the minimax rate of estimation.
-Consequently, if $\wc{f}$ converges at the minimax rate in squared $L^2(P_n)$-norm, then $T_{n,h}\wc{f}$ will converge at the minimax rate in squared $L^2(P)$-norm.
-It follows immediately from Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} (when $s = 1$) or Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} (when $s > 1$) that $T_{n,h}\wh{f}$ achieves the optimal rate of convergence in $L^2(P)$.
-
-\begin{theorem}
-	\label{thm:laplacian_eigenmaps_estimation_out_of_sample}
-	Suppose Model~\ref{def:model_flat_euclidean}.
-There exist constants $c$, $C$, and $N$ that do not depend on $f_0$ or $n$ such that each the following statements hold with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\} - Ch^d\exp\{-cnh^d\}$,  for all $n \geq N$ and for any $\delta \in (0,1)$.
-	\begin{itemize}
-		\item If $f_0 \in H^1(\mc{X};M)$, the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_fo}, and the out-of-sample extension $T_{n,h}\wh{f}$ is computed with bandwidth $h = n^{-1/d}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel}, then
-		\begin{equation*}
-		\|T_{n,h}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2/(2 + d)}.
-		\end{equation*}
-		\item If $f_0 \in H_0^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$ for some $s \in \mathbb{N}, s \geq 2$, and the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_ho}, and the out-of-sample extension $T_{n,h}\wh{f}$ is computed with bandwidth $h = n^{-1/(2(s - 1) + d)}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel} and~\ref{asmp:ho_kernel}, then
-		\begin{equation*}
-		\|T_{n,h}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2s/(2s + d)}.
-		\end{equation*}
-	\end{itemize}
-\end{theorem}
-Some remarks:
-\begin{itemize}
-	\item \emph{Data-driven parameter tuning}.
-Since $T_{n,h}\wh{f}$ is defined out-of-sample, we can use sample splitting or cross-validation methods to tune hyperparameters, which we could not do for the original estimator $\wh{f}$.
-For instance, we can (i) split the sample into two halves, (ii) use the first half to compute $T_{n,h}\wh{f}$ for various values of $\varepsilon$, $h$, and $K$, (iii) choose the optimal values of these three hyperparameters by minimizing error on the held out set.
-Practically speaking, cross-validation is one of the most common approaches to choosing hyperparameters.
-Theoretically, it is known that choosing hyper-parameters through sample splitting can result in estimators that optimally adapt to the order of regularity $s$ \citep{gyorfi2006}.
-In other words, it leads to estimators that are rate-optimal (up to $\log n$ factors), even when $s$ is unknown.
-We believe similar arguments should imply that $T_{n,h}\wh{f}$ is adaptive in this sense when $\varepsilon,h$ and $K$ are chosen by sample splitting, but the details should be worked out carefully.
-
-	% (AG 8/31/21): For now, I punted on proving formal results regarding sample-splitting. I can return later during a revision. 
-	\item \emph{Alternatives to extension by kernel smoothing}.
-There exist other approaches to extending a function $f$ from its evaluations $\{f(X_1),\ldots,f(X_n)\}$: for instance, minimum-norm interpolation in some RKHS or Banach space or 1-nearest neighbors regression.
-We consider extension by kernel smoothing because it is a simple and statistically optimal procedure that does not require any knowledge of the domain $\mc{X}$ or distribution $P$---as we have argued, this latter property is one of the main selling points of Laplacian eigenmaps as a tool for nonparametric regression.
-One potential drawback to kernel smoothing is that it can change the value of the estimate at data, meaning $\wh{f} \neq (\wc{f}(X_1),\ldots, \wc{f}(X_n))$.
-More sophisticated procedures based on Nystr\"{o}m extension of the eigenvectors $\{v_1,\ldots,v_K\}$~\citep{fowlkes2004} inherit the generality of kernel smoothing while resulting in genuine extrapolation; however, the statistical properties of these procedures are substantially more difficult to analyze.
-\end{itemize}
-
 \section{Experiments}
@@
 
-In this section we empirically demonstrate that Laplacian eigenmaps is a reasonably good alternative to population-level spectral projection, even at moderately large sample sizes $n$.
+In this section we empirically demonstrate that the PCR-LE estimator and test are reasonably good alternatives to population-level spectral series methods, even at moderate sample sizes $n$.
 In order to compare the two approaches, in our experiments we stick to the simple case where the design distribution $P$ is the uniform distribution over $\mc{X} = [-1,1]^d$, and we have simple closed-form expressions for the eigenfunctions of $\Delta_P$.
-In general, it is not easy to analytically compute these eigenfunctions, which is part of the appeal of Laplacian eigenmaps.
-
-\begin{figure*}[tb]
+In general, it is not easy to analytically compute these eigenfunctions, which is part of the appeal of LE and PCR-LE.
+
+\begin{figure*}[b]
 	\includegraphics[width=.245\textwidth]{../figures/mse/mse_by_sample_size_1d_1s.pdf}
@@
 	\includegraphics[width=.245\textwidth]{../figures/mse/mse_by_sample_size_2d_2s.pdf}
-	\includegraphics[width=.245\textwidth]{../figures/out_of_sample_mse/mse_by_sample_size_1d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{../figures/out_of_sample_mse/mse_by_sample_size_1d_2s.pdf}
-	\includegraphics[width=.245\textwidth]{../figures/out_of_sample_mse/mse_by_sample_size_2d_1s.pdf}
-	\includegraphics[width=.245\textwidth]{../figures/out_of_sample_mse/mse_by_sample_size_2d_2s.pdf}
-	\caption{Mean squared error (mse) of Laplacian eigenmaps and spectral projection estimators.
-Top row: in-sample mse of Laplacian eigenmaps (\texttt{LE}) and a spectral projection estimator (\texttt{SP}) as a function of sample size $n$.
-Bottom row: out-of-sample mse of Laplacian eigenmaps plus kernel smoothing (\texttt{LE+KS}) and a spectral projection estimator.
+	\caption{In-sample mean squared error (mse) of PCR-LE (\texttt{LE}) vs.
+population-level spectral series (\texttt{SS}) estimator, as a function of sample size $n$.
 Each plot is on the log-log scale, and the results are averaged over 400 repetitions.
-All estimators are tuned for optimal average mse.
+All estimators are tuned for optimal average mse, separately at each value of $n$.
 The black line shows the minimax rate (in slope only; the intercept is chosen to match the observed error).}
@@
 \paragraph{Estimation.}
-In our first experiment, we compare the mean-squared error of the Laplacian eigenmaps estimator $\wh{f}$  to that of its population-level counterpart $\wt{f}$.
-We vary the sample size from $n = 1000$ to $n = 4000$; sample $n$ design points $\{X_1,\ldots,X_n\}$ from the uniform distribution on the cube $[-1,1]^d$; and sample responses $Y_i$ according to~\eqref{eqn:model} with regression function $f_0 = M/\lambda_K^{s/2} \cdot \psi_K$ for $K \asymp n^{d/(2s + d)}$ (the pre-factor $M/\lambda_K^{s/2}$ is chosen so that $|f_0|_{H^s(\mc{X})}^2 = M^2$).
+In our first experiment, we compare the mean-squared error of the PCR-LE estimator $\wh{f}$  to that of its population-level counterpart $\wt{f}$.
+We vary the sample size from $n = 1000$ to $n = 4000$; sample $n$ design points $\{X_1,\ldots,X_n\}$ from the uniform distribution on the cube $[-1,1]^d$; and sample responses $Y_i$ according to~\eqref{eqn:model} with regression function $f_0 = M/\rho_K^{s/2} \cdot \psi_K$ for $K \asymp n^{d/(2s + d)}$ (the pre-factor $M/\rho_K^{s/2}$ is chosen so that $|f_0|_{H^s(\mc{X})}^2 = M^2$).
 In Figure~\ref{fig:fig1} we show the in-sample mean-squared error of the two estimators as a function of $n$, for different dimensions $d$ and order of smoothness $s$.
-We see that all estimators have mean-squared error converging to zero at roughly the minimax rate.
-While unsurprisingly population-level spectral projection has smaller error, generally speaking the error of Laplacian eigenmaps approaches that of population-level spectral projection as $n$ gets larger.
-
-We also compare the out-of-sample estimation error, over a held-out test set, of Laplacian eigenmaps plus kernel smoothing to the spectral projection estimator.
-The story is much the same, supporting our theoretical claim that the additional error incurred by kernel smoothing of Laplacian eigenmaps is negligible.
+We see that both estimators have mean-squared error converging to zero at roughly the minimax rate.
+While the unsurprisingly population-level spectral series estimator has the smaller error, generally speaking the error of PCR-LE approaches that of the population-level spectral series method as $n$ gets larger.
 
 \paragraph{Testing.} 
-In our second experiment, we compare the Laplacian eigenmaps test $\varphi$ against the  population-level spectral projection test $\wt{\varphi}$.
+In our second experiment, we compare the PCR-LE test $\varphi$ against the  population-level spectral series test $\wt{\varphi}$.
 The setup is generally the same as that of our first experiment, but to get an empirical estimate of the critical radius the details are necessarily somewhat more complicated.
-First we take $\mc{F} = \{M/\lambda_k^{s/2} \psi_k\}_{k = 1}^{n}$ to be a discrete subset of $H^1(\mc{X};M)$.
-Then, for each $f_0 \in \mc{F}$, we run a given test $\phi$ (either the Laplacian eigenmaps test $\phi = \varphi$, or the population-level spectral projection test $\phi = \wt{\varphi}$.) and record whether it was a false negative or true positive.
+First we take $\mc{F} = \{M/\rho_k^{s/2} \psi_k\}_{k = 1}^{n}$ to be a discrete subset of $H^1(\mc{X};M)$.
+Then, for each $f_0 \in \mc{F}$, we run a given test $\phi$ (either the PCR-LE test $\phi = \varphi$, or the population-level spectral series test $\phi = \wt{\varphi}$) and record whether it was a false negative or true positive.
 We repeat this process over $100$ replications, giving a Monte Carlo estimate of the type II error $E_{f_0}[1 - \phi]$ for each $f_0 \in \mc{F}$.
@@
 
-In Figure~\ref{fig:fig2}, we see that the estimated critical radii of both Laplacian eigenmaps and population-level spectral projection tests are quite close to each other, and converge at roughly the minimax rate.
-\begin{figure*}[tb]
+In Figure~\ref{fig:fig2}, we see that the estimated critical radii of both the PCR-LE and population-level spectral series tests are quite close to each other, and converge at roughly the minimax rate.
+\begin{figure*}[b]
 	\includegraphics[width=.245\textwidth]{../figures/testing/eigenfunction/critical_radius_by_sample_size_1d_1s.pdf}
@@
 	\includegraphics[width=.245\textwidth]{../figures/testing/eigenfunction/critical_radius_by_sample_size_2d_2s.pdf}
-	\caption{Worst-case testing risk Laplacian eigenmaps (\texttt{LE}) and spectral projection (\texttt{SP}) tests, as a function of sample size $n$.
+	\caption{Worst-case testing risk for PCR-LE (\texttt{LE}) and spectral series (\texttt{SP}) tests, as a function of sample size $n$.
 Plots are on the same scale as Figure~\ref{fig:fig1}, and black line shows the minimax rate.
@@
 \paragraph{Tuning parameters.}
-Our first two experiments demonstrate that Laplacian eigenmaps methods have comparable statistical performance to spectral projection methods.
-Laplacian eigenmaps depends on two tuning parameters, and in our final experiment we investigate the importance of both, focusing now on estimation.
-In Figure~\ref{fig:fig3}, we see how the mean-squared error of Laplacian eigenmaps changes as each tuning parameter is varied.
+Our first two experiments demonstrate that PCR-LE methods have comparable statistical performance to population-leve spectral series methods.
+PCR-LE depends on two tuning parameters, and in our final experiment we investigate the importance of both, focusing now on estimation.
+In Figure~\ref{fig:fig3}, we see how the mean-squared error of PCR-LE changes as each tuning parameter is varied.
 As suggested by our theory, properly choosing the number of eigenvectors $K$ is crucial: the mean-squared error curves, as a function of $K$, always have a sharply defined minimum.
@@
 This squares completely with our theory, which requires that the number of eigenvectors $K$ be much more carefully tuned that the graph radius $\varepsilon$.
-
-We also plot the out-of-sample mean squared error of Laplacian eigenmaps plus kernel smoothing, as a function of its various tuning parameters (which include the bandwidth $h$ as well as $\varepsilon$ and $K$.) Here the relationship between theory and empirics is more nuanced.
-On the one hand, empirically it seems that the optimal choice of bandwidth parameter $h$ is usually smaller than $\varepsilon$, as suggested by our theory.
-On the other hand, for Laplacian eigenmaps plus kernel smoothing we see that as a function of $K$, the mean-squared error curves are close to their minima even when we choose many more eigenvectors than is optimal for Laplacian eigenmaps.
-This is not reflected in our theory, which requires that $K$ be chosen in the same tight range as was required for Laplacian eigenmaps to be optimal in-sample.
-However, it does make intuitive sense: extension by kernel smoothing further attenuates the noise, making the algorithm more forgiving to overfitting during the Laplacian eigenmaps step.
 
@@
 	\includegraphics[width=.245\textwidth]{../figures/tuning/sobolev/mse_by_radius_2d_1s.pdf}  
-	\caption{Mean squared error of Laplacian eigenmaps (\textcolor{red}{red}), Laplacian eigenmaps plus kernel smoothing (\textcolor{blue}{blue}), and spectral projection (\textcolor{green}{green}) as a function of tuning parameters.
+	\caption{Mean squared error of PCR-LE (\textcolor{red}{red}), and population-level spectral series (\textcolor{green}{green}) estimators as a function of tuning parameters.
 Top row: the same regression function $f_0$ as used in Figure~\ref{fig:fig1}.
-Bottom row: the regression function $f_0 \propto \sum_{k} 1/\lambda_k^{1/2} \psi_k$.
+Bottom row: the regression function $f_0 \propto \sum_{k} 1/\rho_k^{1/2} \psi_k$.
 For all experiments, the sample size $n = 1000$, and the results are averaged over $200$ repetitions.
-In each panel, all tuning parameters except the one being varied are set to their optimal values.
-For Laplacian eigenmaps plus kernel smoothing, circular points and a solid line are used to denote the error as a function of the graph radius $\varepsilon$, whereas triangular points and a dashed line are used to denote the error as a function of the bandwidth $h$.}
+In each panel, all tuning parameters except the one being varied are set to their optimal values.}
 	\label{fig:fig3}
@@
 
-In this work, we have derived upper bounds on the rates of convergence for methods based on Laplacian eigenmaps, which imply that in various settings these methods are minimax rate-optimal over Sobolev classes.
+In this work, we have derived upper bounds on the rates of convergence for regression with PCR-LE, which imply that in various settings the PCR-LE estimator and test are minimax rate-optimal over Sobolev classes.
 Importantly, these upper bounds hold under nonparametric conditions on the design density $p$, and allow for $p$ to be unknown and, potentially, supported on a low-dimensional manifold.
 Our results help explain the practical success of methods which leverage graph Laplacian eigenvectors for regression.
-They also distinguish Laplacian eigenmaps from more traditional spectral projection procedures, which rely on a density-dependent basis and thus require the density be known a priori.
+They also distinguish such methods from more traditional spectral series procedures, which rely on a density-dependent basis and thus require the density be known a priori.
 
 Of course, there do exist other methods for nonparametric regression which achieve optimal rates of convergence under similar (or indeed weaker) conditions on $p$.
-These include other graph-based approaches---graph Laplacian regularization---methods besides spectral projection---e.g.
+These include other graph-based approaches---graph Laplacian regularization---methods besides spectral series methods---e.g.
 kernel smoothing, local polynomial regression, thin-plate splines---and continuum spectral projection methods which use the eigenfunctions of an operator defined independently of $p$.
-To be clear, we do not advocate Laplacian eigenmaps over these alternatives.
-Rather, we view our results as theoretically justifying a place for Laplacian eigenmaps in the nonparametric regression toolbox.
-
-That being said, Laplacian eigenmaps does have certain advantages over each of the aforementioned approaches.
-We now conclude by outlining some of these advantages (limiting our discussion to estimation):
+To be clear, we do not advocate PCR-LE over these alternatives.
+Rather, we view our results as theoretically justifying a place for regression using Laplacian Eigenmaps in the nonparametric regression toolbox.
+
+That being said, PCR-LE does have certain advantages over each of the aforementioned approaches. We now conclude by outlining some of these advantages (limiting our discussion to estimation):
 \begin{itemize}
 	\item \emph{Optimality over high-dimensional Sobolev spaces}.
-Graph Laplacian regularization achieves minimax optimal rates over $H^1(\mc{X})$ only when $d \in \{1,2,3,4\}$ \citep{sadhanala16, green2021}; when $d \geq 5$, graph Laplacian regularization appears to be suboptimal.
-In contrast, Laplacian eigenmaps is optimal over $H^1(\mc{X})$ for all dimensions $d$.
-An even more stark contrast can be drawn between Laplacian eigenmaps (and other spectral projection estimators) and the first order thin-plate spline, the solution of the variational problem
+As mentioned in the introduction, graph Laplacian regularization provably achieves minimax optimal rates over $H^1(\mc{X})$ only when $d \in \{1,2,3,4\}$ \citep{sadhanala16, green2021}; when $d \geq 5$, graph Laplacian regularization appears to be suboptimal.
+In contrast, PCR-LE is optimal over $H^1(\mc{X})$ for all dimensions $d$.
+An even more stark contrast can be drawn between PCR-LE (and other spectral series estimators) and the first order thin-plate spline, the solution of the variational problem
 	\begin{equation}
@@
 	\item \emph{Manifold adaptivity}.
-An oft-recommended alternative to the population-level spectral projection methods considered in Section~\ref{subsec:spectral_projection} is to run ordinary least-squares regression using eigenfunctions of a density-independent operator, such as the Laplacian $\Delta = \sum_{i = 1}^{d} \partial^2f/\partial x_i^2$.
+An oft-recommended alternative to the population-level spectral series method considered in Section~\ref{subsec:spectral_projection} is to run OLS using eigenfunctions of a density-independent differential operator, such as the unweighted Laplacian $\Delta = \sum_{i = 1}^{d} \partial^2f/\partial x_i^2$.
 Under the conditions of Model~\ref{def:model_flat_euclidean}, such a method will indeed be rate-optimal, though the upper bounds may come with undesirably large constants if $p$ is very non-uniform.
-However, in contrast with Laplacian eigenmaps, under Model~\ref{def:model_manifold} this method cannot achieve the faster minimax rates of convergence.
+However, in contrast with PCR-LE, under Model~\ref{def:model_manifold} this method cannot achieve the faster minimax rates of convergence.
 	\item \emph{Density adaptivity}.
-It is natural to ask whether the two stage estimator $T_{n,h}\wh{f}$ defined in Section~\ref{sec:out_of_sample} has any advantage over the simpler approach of directly kernel smoothing the responses, i.e.
-using the estimator $T_{n,h}{\bf Y}$ (possibly for a different choice of $h$).
-In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we answer this question in the affirmative, by giving a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that $\Ebb\|f_0 - T_{n,h}\wh{f}\|_P^2$ is of a strictly lower order than $\inf_{h'} \Ebb\|f_0 - T_{n,h'}{\bf Y}\|_P^2$.
-This is possible because Laplacian eigenmaps induces a completely different bias than kernel smoothing.
-For example, when $f_0$ and $p$ satisfy the so-called \emph{cluster assumption}---meaning $f_0$ is piecewise constant in high-density regions (clusters) of $p$---then the bias of Laplacian eigenmaps can be much smaller than that of kernel smoothing (for equivalent levels of variance).
-
-	We emphasize that this does not contradict the well-known optimality properties of kernel smoothing over e.g.
-H\"{o}lder balls.
-Rather, in the standard nonparametric regression setup---which we adopt in the main part of this paper, and in which $P$ is assumed to be equivalent to Lebesgue measure---the biases of Laplacian eigenmaps and kernel smoothing happen to be equivalent.
+In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we give a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that the expected in-sample mean squared error of PCR-LE is smaller than that of either kernel smoothing or least squares using eigenfunctions of $\Delta$.
+This is possible because PCR-LE induces a completely different bias than these latter two methods.
+In particular, when $f_0$ and $p$ satisfy the so-called \emph{cluster assumption}---meaning $f_0$ is piecewise constant in high-density regions (clusters) of $p$---then the bias of PCR-LE can be much smaller (for equivalent levels of variance) than that of kernel smoothing or least-squares with eigenfunctions of $\Delta$.
+
+	We emphasize that this does not contradict the well-known optimality properties of, for example, kernel smoothing over H\"{o}lder balls.
+Rather, in the standard nonparametric regression setup---which we adopt in the main part of this paper, and in which $P$ is assumed to be equivalent to Lebesgue measure---the biases of PCR-LE and kernel smoothing happen to be equivalent.
 But when $P$ is sufficiently non-uniform, this is no longer the case.

### 2021-10-16 code/plot_demos/eigenmaps_mse.R
Updated code for producing plots for Laplacian eigenmaps.

@@ -10,9 +10,9 @@ source("plot_methods.R")
 source("sample.R")
 
 # User entered information.
-data_directory <- "data/thesis/mse/eigenfunction_1s_1d"
-# plot_directory <- "plots/thesis/mse/eigenfunction/"
-plot_directory <- "../../thesis/talk/figures/mse" # Please change this to whichever directory you prefer.
+data_directory <- "data/laplacian_eigenmaps/mse/eigenfunction_2s_2d"
+plot_directory <- "plots/laplacian_eigenmaps/mse/eigenfunction/"
+# plot_directory <- "../../thesis/talk/figures/mse" # Please change this to whichever directory you prefer.
 if(!exists(plot_directory)){dir.create(plot_directory)}
 
 load(file.path(data_directory,"configs.R"))
@@ -27,18 +27,18 @@ load(file.path(data_directory,"Ys.R"))
 
 # Subset data to methods you actually want to plot.
 plot_methods <- c("laplacian_eigenmaps",
-                  "laplacian_smoothing")
+                  "spectral_projection")
 mse <- lapply(mse,FUN = function(m){m[names(methods) %in% plot_methods]})
 methods <- methods[names(methods) %in% plot_methods]
 
 ## Plot 1: Mean squared error as a function of n.
 plot_name <- paste0("mse_by_sample_size_",d,"d_",s,"s.pdf")
 pdf(file.path(plot_directory,plot_name))
-par(mar = c(5.1,6,4.1,4.1),lwd = 3, cex.main = 4) # Prevent the left hand side from getting cut off.
-title <- paste0("d = ", d,", s = ",s,".")
-colors <-  c("purple","deepskyblue2")
-plot_methods_abbr <- c("LE","LS") # for the legend
-plot_best_mse(methods,mse,title = title, sd = F,rate = F,legend = F, cols = colors)
-legend("bottomleft", legend = plot_methods_abbr, col = colors, pch = 20, lwd = par()$lwd,
-       bg = "white", inset = .05, cex = 1.75)
+par(mar = c(5.1,6,4.1,4.1), cex.main = 2.5) # Prevent the left hand side from getting cut off.
+title <- paste0("d = ", d,", s = ",s,". Minimax slope = -",2*s,"/",2*s + d,".")
+colors <-  c("red","green")
+plot_methods_abbr <- c("LE","SS") # for the legend
[…]

### 2021-10-16 code/plot_demos/eigenmaps_tuning_parameters.R

@@ -9,12 +9,9 @@ source("plot_methods.R")
 source("misc.R")
 
 # User entered information.
-data_directory <- "data/thesis/mse/eigenfunction_1s_5d_2"
-plot_directory <- "plots/thesis/mse/eigenfunction" # Please change this to whichever directory you prefer.
+data_directory <- "data/laplacian_eigenmaps/tuning/sobolev_1s_2d"
+plot_directory <- "plots/laplacian_eigenmaps/tuning/sobolev" # Please change this to whichever directory you prefer.
 if(!exists(plot_directory)){dir.create(plot_directory)}
-
-plot_n <- 4000                                  # Which sample size would you like to use?
-plot_methods <- c("laplacian_eigenmaps")
 
 # Load data
 load(file.path(data_directory,"configs.R"))
@@ -27,10 +24,16 @@ load(file.path(data_directory,"Xs.R"))
 load(file.path(data_directory,"f0s.R"))
 load(file.path(data_directory,"Ys.R"))
 
+# Subset data to methods you actually want to plot.
+plot_n <- 1000
+plot_methods <- c("laplacian_eigenmaps",
+                  "spectral_projection")
 stopifnot(plot_n %in% ns)
+stopifnot(all(plot_methods %in% names(methods)))
 mse <- mse[ns %in% plot_n][[1]][names(methods) %in% plot_methods]
 if(exists("test_mse")) test_mse <- test_mse[ns %in% plot_n][[1]][names(methods) %in% plot_methods]
 thetas <- thetas[ns %in% plot_n][[1]][names(methods) %in% plot_methods]
+methods <- methods[names(methods) %in% plot_methods]
 
 # Plotting parameters for all plots
 # Capitalization function from https://rstudio-pubs-static.s3.amazonaws.com/408658_512da947714740b99253228f084a08a9.html.
@@ -75,8 +78,8 @@ for(ii in alg_indx)
 # Plotting parameters
 xlims <- c(min(Ks),max(Ks))
 ylims <- c(min(plot_mse),max(plot_mse))
-cols <- c("red","blue","green")
-stopifnot(ncol(plot_mse) <= 3)
+cols <- c("red","green")
[…]

### 2021-10-16 code/plot_demos/testing.R

@@ -10,8 +10,8 @@ source("plot_methods.R")
 source("sample.R")
 
 # User entered information.
-data_directory <- "data/laplacian_eigenmaps/testing/20210601064335"
-plot_directory <- "plots/laplacian_eigenmaps/testing/experimental" # Please change this to whichever directory you prefer.
+data_directory <- "data/laplacian_eigenmaps/testing/eigenfunction_2d_2s"
+plot_directory <- "plots/laplacian_eigenmaps/testing/eigenfunction" # Please change this to whichever directory you prefer.
 if(!exists(plot_directory)){dir.create(plot_directory)}
 
 # Plotting parameters
@@ -28,7 +28,6 @@ load(file.path(data_directory,"power.R"))
 
 # Subset data to methods you actually want to plot.
 plot_methods <- c("laplacian_eigenmaps",
-                  "laplacian_eigenmaps_plus_kernel_smoothing",
                   "spectral_projection")
 err <- lapply(err,FUN = function(m){m[names(methods) %in% plot_methods]})
 power <- lapply(power,FUN = function(m){m[names(methods) %in% plot_methods]})
@@ -37,7 +36,7 @@ methods <- methods[names(methods) %in% plot_methods]
 # Plot critical radius.
 plot_name <- paste0("critical_radius_by_sample_size_",d,"d_",s,"s.pdf")
 pdf(file.path(plot_directory,plot_name))
-par(mar = c(5.1,6,4.1,4.1)) # Prevent the left hand side from getting cut off.
+par(mar = c(5.1,6,4.1,4.1),cex.main = 2.5) # Prevent the left hand side from getting cut off.
 plot_testing_critical_radius(methods,err,type_II_error = runif(100,0,1),
                              sd, cols, verbose = T)
 dev.off()

### 2021-10-16 code/plot_methods.R

@@ -66,7 +66,7 @@ plot_best_mse <- function(methods,mse,sd = T,validate = validate_mse,cols = NULL
     names(plot_dfs_best_mse)[[jj]] <- case_when(
       name == "laplacian_smoothing" ~ "LS",
       name == "laplacian_eigenmaps" ~ "LE",
-      name == "spectral_projection" ~ "SP",
+      name == "spectral_projection" ~ "SS",
       name == "least_squares"       ~ "LS",
       name == "laplacian_eigenmaps_plus_kernel_smoothing" ~ "LE+KS"
     ) 
@@ -157,7 +157,7 @@ plot_testing_critical_radius <- function(methods,err,type_II_error = .5,
     abbr <- case_when(
       name == "laplacian_smoothing" ~ "LS",
       name == "laplacian_eigenmaps" ~ "LE",
-      name == "spectral_projection" ~ "SP",
+      name == "spectral_projection" ~ "SS",
       name == "least_squares"       ~ "LS",
       name == "laplacian_eigenmaps_plus_kernel_smoothing" ~ "LE+KS"
     ) 

### 2021-10-25 manuscripts/laplacian_eigenmaps/appendix.tex
Completed revision of intro in response to Ryan's comments.

@@
 \end{equation}
-Since the eigenfunctions $\{\psi_k\}$ form an orthonormal basis of $L^2(P)$, and $f_0 \in \mc{H}^{s}(\mc{X}) \subseteq L^2(P)$, we can write the squared bias in terms of squared Fourier coefficients of $f_0$, leading to the following upper bound,
+Since the eigenfunctions $\{\psi_k\}$ form an orthonormal basis of $L^2(\mc{X})$ (with respect to the inner-product $\dotp{\cdot}{\cdot}_P$) and $f_0 \in \mc{H}^{s}(\mc{X}) \subseteq L^2(\mc{X})$, we can write the squared bias in terms of squared Fourier coefficients of $f_0$, leading to the following upper bound,
 \begin{equation*}
@@
 	\end{equation*}
-	In the above, the last equality relies on the fact that $v_k$ are orthonormal in $L^2(P_n)$. Using the fact that the eigenvalues are in increasing order, we obtain
+	In the above, the last equality relies on the fact that $v_k$ are orthonormal with respect to $\dotp{\cdot}{\cdot}_n$. Using the fact that the eigenvalues are in increasing order, we obtain
 	\begin{equation*}
@@
 We will divide our analysis in two cases, depending on whether $s$ is odd or even, but before we do this we state some facts that will be applicable to both cases.
-First, we  recall that $L_{P,\varepsilon}$ is self-adjoint in $L^2(P)$, meaning $\dotp{L_{P,\varepsilon}f}{g}_{P} = \dotp{f}{L_{P,\varepsilon}g}_{P}$ for all $f, g \in L^2(P)$.
+First, we  recall that $L_{P,\varepsilon}$ is self-adjoint in $L^2(P)$, meaning $\dotp{L_{P,\varepsilon}f}{g}_{P} = \dotp{f}{L_{P,\varepsilon}g}_{P}$ for all $f, g \in L^2(\mc{X})$.
 We also recall the definition of the Dirichlet energy $E_{P,\varepsilon}(f;\mc{X})$,
@@
 with probability at least $1 - \exp(-K)$.
-Consequently $\wt{f}$ has $L^2(P_n)$-error of at most $\sigma^{2s}$ times our upper bound on the error of $\wh{f}$, and for any choice of $\sigma$ that is constant in $n$ the estimator $\wc{f}$ will also be rate-optimal.
+Consequently $\|\wt{f} - f_0\|_n^2$is at most $\sigma^{2s} \cdot \|\wh{f} - f_0\|_n^2$, and for any choice of $\sigma$ that is constant in $n$ the estimator $\wc{f}$ will also be rate-optimal.
 

### 2021-10-25 manuscripts/laplacian_eigenmaps/arxiv/laplacian_eigenmaps_regression.tex
Completed revision of intro in response to Ryan's comments.

@@
 % Title 
-\begin{center} {\Large{\bf{Minimax-optimal Regression over Sobolev Spaces \\
+\begin{center} {\Large{\bf{Minimax Optimal Regression over Sobolev Spaces \\
 				\vspace{.2cm}
-				via Laplacian Eigenmaps with Neighborhood Graphs} \\}
-			\textcolor{red}{(TODO): Keep this title?}}
+				via Laplacian Eigenmaps on Neighborhood Graphs}}}
 
@@
 PCR-LE works by projecting a vector of observed responses ${\bf Y} = (Y_1,\ldots,Y_n)$ onto a subspace spanned by certain eigenvectors of a neighborhood graph Laplacian.
-We show that in many cases PCR-LE achieves minimax rates of convergence over continuum function classes.
-Specifically, we consider the problem of random design regression over Sobolev spaces $H^s(\mc{X})$---where $\mc{X} \subseteq \Rd$ is the support of the design distribution $P$---and show that under sufficient smoothness conditions on the design density $p$, PCR-LE achieves the optimal rates for both estimation (where the optimal rate is known to be $n^{-2s/(2s + d)}$) and goodness-of-fit testing ($n^{-4s/(4s + d)}$).
+We show that PCR-LE achieves minimax rates of convergence over continuum function classes.
+Specifically, we consider the problem of random design regression over Sobolev spaces $H^s(\mc{X})$---where $\mc{X} \subseteq \Rd$ is the support of the design distribution $P$---and show that under sufficient smoothness conditions on the design density $p$, PCR-LE achieves the optimal rates for both estimation (where the optimal rate in squared $L^2$ norm is known to be $n^{-2s/(2s + d)}$) and goodness-of-fit testing ($n^{-4s/(4s + d)}$).
 We also show that PCR-LE is \emph{manifold adaptive}: that is, we consider the situation where $\mc{X}$ is a manifold of small intrinsic dimension $m$, and give upper bounds establishing that PCR-LE achieves the faster minimax estimation ($n^{-2s/(2s + m)}$) and testing ($n^{-4s/(4s + m)}$) rates of convergence.
-Interestingly, these rates are almost always much faster than the known rates of convergence of graph Laplacian eigenvectors to their population-level limits; in other words, for this problem regression with learned features appears to be much easier, statistically speaking, than feature learning.
+Interestingly, these rates are almost always much faster than the known rates of convergence of graph Laplacian eigenvectors to their population-level limits; in other words, for this problem regression with learned features appears to be much easier, statistically speaking, than feature learning itself.
 We support these theoretical results with empirical evidence.

### 2021-10-25 manuscripts/laplacian_eigenmaps/main_text.tex
Completed revision of intro in response to Ryan's comments.

@@
 Laplacian Eigenmaps (LE)~\citep{belkin03a} is a method for nonlinear dimensionality reduction and data representation.
-Given scattered data points $\{X_1,\ldots,X_n\} \subset \Reals^d$, LE maps each $X_i$ to a vector $(v_{i,1},\ldots,v_{i,K})$ according to the following steps.
+Given data points $\{X_1,\ldots,X_n\} \subset \Reals^d$, LE maps each $X_i$ to a vector $(v_{1,i},\ldots,v_{K,i})$ according to the following steps.
 \begin{enumerate}
@@
 \end{enumerate} 
-A natural way to use LE is by taking the vectors $\{(v_{i,1},\ldots,v_{i,K})\}_{i = 1}^{n}$ to be features in a downstream supervised learning algorithm.
+A natural way to use LE is by taking the vectors $\{(v_{1,i},\ldots,v_{K,i})\}_{i = 1}^{n}$ to be features in a downstream supervised learning algorithm.
 In this paper, we study a simple method along these lines: Principal Components Regression with Laplacian-Eigenmaps (PCR-LE), a method for nonparametric regression which operates by running ordinary least squares (OLS) using the features learned by LE.
@@
 Loosely speaking, these works model the design points as being independent samples from a distribution $P$ with density $p$, and show that in this case the eigenvectors of the graph Laplacian $L$ are good empirical approximations of population-level objects.
-These population-level objects are the eigenfunctions $\psi_k$---meaning solutions, along with eigenvalues $\rho_k$, to the equation $\Delta_P \psi_k = \rho_k \psi_k$--- of the density-weighted Laplacian operator
+These population-level objects are eigenfunctions $\psi_k$---meaning solutions, along with eigenvalues $\rho_k$, to the equation $\Delta_P \psi_k = \rho_k \psi_k$--- of a density-weighted Laplacian operator defined via:
 \begin{equation}
@@
 These aforementioned works explain what kind of representation LE learns, and the accuracy with which it learns this representation.
-However, this theory does not address the convergence rates of PCR-LE.
-That is the major question we answer in this paper.
-We adopt the classical model of nonparametric regression with random design, where one observes independent pairs $(X_1,Y_1),\ldots,(X_n,Y_n)$ of design points and responses.
+However, this theory does not focus on the statistical properties of PCR-LE for classical regression problems such as estimation and testing.
+That is the major question we address in this paper.
+We adopt the usual model of nonparametric regression with random design, where one observes independent pairs $(X_1,Y_1),\ldots,(X_n,Y_n)$ of design points and responses.
 We assume the design points $\{X_1,\ldots,X_n\}$ are sampled from an unknown distribution $P$ supported on $\mc{X} \subseteq \Rd$, and the responses follow a signal plus Gaussian noise model,
@@
 \end{equation}
-with noise variables $w_i$ independent of $X_i$.
+with noise variables $w_i$  independent of design points $X_i$.
 The task is to learn the regression function $f_0$, which is unknown but assumed to belong to a Sobolev space $H^s(\mc{X})$.
@@
 At first glance, the optimality of PCR-LE is surprising.
-As a function of the sample size $n$, known rates of convergence of LE are much slower than the minimax-optimal rates of convergence over $H^s(\mc{X})$ (roughly speaking $n^{-2s/(2s + d)}$ for estimation and $n^{-4s/(4s + d)}$ for testing).
+As a function of the sample size $n$, known rates of convergence of LE are much slower than the minimax optimal rates of convergence over $H^s(\mc{X})$ (roughly speaking $n^{-2s/(2s + d)}$ for estimation and $n^{-4s/(4s + d)}$ for testing).
 This is not due to suboptimal theory for LE.
-Rather, it reflects a fundamental fact: for this problem, regression using learned features does not rely on accurately learning the features.
+Rather, it reflects a fundamental fact: \emph{for this problem, regression using learned features does not rely on accurately learning the features}.
 It is essential to employ modes of analysis which exploit this fact in order to derive sharp rates of convergence for PCR-LE.
@@
 We study regression over Sobolev spaces in part because, generally speaking, the minimax rates are well-understood; as mentioned before, when the domain $\mc{X}$ is full-dimensional they are $n^{-2s/(2s + d)}$ for estimation, and $n^{-4s/(4s + d)}$ for testing.
-For this reason, regression over Sobolev spaces is a good setting in which to see whether PCR-LE measures up to simpler and more classical minimax-optimal approaches, which have strong theoretical guarantees but are less often used in practice.
+For this reason, regression over Sobolev spaces is a good setting in which to see whether PCR-LE measures up to more standard minimax optimal approaches, which have strong theoretical guarantees but are less often used in practice because their applicability is more limited.
+\textcolor{red}{(TODO): This last clause still seems weak to me.
+Does Siva agree?}
 
@@
 \end{equation} 
-Under appropriate boundary conditions, the Sobolev spaces $H^s(\mc{X})$ can roughly be characterized as consisting of those functions $f = \sum_{k} a_k \psi_k \in L^2(\mc{X})$ for which the generalized Fourier coefficients $\{a_k\}_{k = 1}^{\infty}$ satisfy the decay condition $\sum_{k} a_k \rho_k^2 < \infty$ (See Section~\ref{subsec:spectral_projection} for more details).
+Under appropriate boundary conditions, the Sobolev spaces $H^s(\mc{X})$ consists of function $f = \sum_{k} a_k \psi_k \in L^2(\mc{X})$ with generalized Fourier coefficients $\{a_k\}_{k = 1}^{\infty}$ that satisfy the decay condition $\sum_{k} a_k \rho_k^2 < \infty$ (See Section~\ref{subsec:spectral_projection} for more details).
 This decay condition justifies the truncated series estimator \eqref{eqn:population-level_spectral_series}, since it means the truncation will incur only a limited amount of bias for any $f_0 \in H^s(\mc{X})$.
-For this reason spectral series regression over Sobolev spaces has been well-studied since at least~\citet{rice1984}\footnote{And of course proposed much earlier in the context of density estimation by \citet{cencov1962}.}, and its optimality properties are by this point well-understood.
+For this reason spectral series regression over Sobolev spaces has been well-studied since at least \citet{rice1984},\footnote{And proposed much earlier in the context of density estimation by \citet{cencov1962}.} and its optimality properties are by this point well-understood.
 
 PCR-LE serves as an empirical approximation to spectral series regression, since the eigenvectors $v_k$ are empirical approximations to the eigenfunctions $\psi_k$.
-Viewed in this light, a major advantage of PCR-LE is that it operates independently of the design distribution $P$.
-In the contrast, the spectral series estimator defined in~\eqref{eqn:population-level_spectral_series} relies on diagonalizing the density-weighted Laplacian $\Delta_P$.
-In our context, where $P$ is treated as unknown,~\eqref{eqn:population-level_spectral_series} must be viewed as an oracle method, and to emphasize this  we henceforth refer to~\eqref{eqn:population-level_spectral_series} as \emph{population-level spectral series regression}.
-On the other hand, naturally PCR-LE incurs some extra error by using an empirical approximation to the underlying basis $\{\psi_k\}_{k = 1}^{\infty}$, and our work shows that in many cases, this extra error is not enough to change the overall rate of convergence.
+Viewed in this light, a major advantage of PCR-LE is that it operates without needing knowledge of the design distribution $P$.
+In contrast, the spectral series estimator defined in~\eqref{eqn:population-level_spectral_series} relies on diagonalizing the density-weighted Laplacian $\Delta_P$.
+In our context, where $P$ is treated as unknown,~\eqref{eqn:population-level_spectral_series} must be viewed as an oracle method, and to emphasize this we henceforth refer to it as \emph{population-level spectral series regression}.
+On the other hand, intuitively PCR-LE incurs some extra error by using an empirical approximation to the underlying basis $\{\psi_k\}_{k = 1}^{\infty}$: our work shows that in many cases, this extra error is not enough to change the overall rate of convergence.
 
@@
 \paragraph{Rates of convergence: population-level spectral series regression.}
-As we have already mentioned, the minimax-optimal rates over Sobolev spaces are generally well-known, as are upper bounds for population-level spectral series methods which match these rates.
-However, we could not find precisely stated results applying to our setting, which is somewhat unusual in the following respects.
+As we have already mentioned, the minimax optimal rates over Sobolev spaces are generally well-known, as are upper bounds for population-level spectral series methods which match these rates.
+However, we could not find precisely stated results applying to our setting, which is quite general in the following respects.
 \begin{enumerate}
 	\item We consider Sobolev spaces $H^s(\mc{X})$ for all combinations of $s$ and $d$.
-This includes the subcritical regime where the smoothness parameter $s$ satisfies $s < d/2$ and so $H^s(\mc{X})$ does not continuously embed into the space of continuous functions $C^0(\mc{X})$.
+This includes the subcritical regime where the smoothness parameter $s$ satisfies $s < d/2$; in this regime $H^s(\mc{X})$ does not continuously embed into the space of continuous functions $C^0(\mc{X})$.
 	\item We consider general design distributions $P$, which may satisfy certain regularity conditions but are not limited to being, say, the uniform distribution over $[0,1]^d$. 
@@
 \begin{itemize}
-	\item  Over a ball in the Sobolev space $H^{s}(\mc{X})$, we establish that the PCR-LE estimator $\wh{f}$ has in-sample mean-squared error of at most on the order of $n^{-2s/(2s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d$.
-	\item We show that a test based on the statistic $\|\wh{f}\|_2^2$ has a squared critical radius on the order of $n^{-4s/(4s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d \in \{1,2,3,4\}$.
+	\item  Over a ball in the Sobolev space $H^{s}(\mc{X})$, we establish that the PCR-LE estimator $\wh{f}$ has in-sample mean-squared error on the order of $n^{-2s/(2s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d$ (Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and~\ref{thm:laplacian_eigenmaps_estimation_ho}).
+	\item We show that a test based on the statistic $\|\wh{f}\|_2^2$ has a squared critical radius on the order of $n^{-4s/(4s + d)}$, for any number of derivatives $s \in \mathbb{N}$ and dimension $d \in \{1,2,3,4\}$ (Theroems~\ref{thm:laplacian_eigenmaps_testing_fo} and~\ref{thm:laplacian_eigenmaps_testing_ho}).
 \end{itemize}
@@
 \begin{itemize}
-	\item Over a ball in the Sobolev space $H^{s}(\mc{X})$, the PCR-LE estimator has in-sample mean squared error of at most $n^{-2s/(2s + m)}$, when $s \in \{1,2,3\}$ and for any $m \in \mathbb{N}$. 
-	\item A test based on the statistic $\|\wh{f}\|_2^2$ has a squared critical radius on the order of $n^{-4s/(4s + m)}$, when $s \in \{1,2,3\}$ and $m \in \{1,2,3,4\}$.
+	\item Over a ball in the Sobolev space $H^{s}(\mc{X})$, the PCR-LE estimator $\wh{f}$ has in-sample mean squared error of at most $n^{-2s/(2s + m)}$, when $s \in \{1,2,3\}$ and for any $m \in \mathbb{N}$ (Theorem~\ref{thm:laplacian_eigenmaps_estimation_manifold}).
+	\item A test based on the statistic $\|\wh{f}\|_2^2$ has a squared critical radius on the order of $n^{-4s/(4s + m)}$, when $s \in \{1,2,3\}$ and $m \in \{1,2,3,4\}$ (Theorem~\ref{thm:laplacian_eigenmaps_testing_manifold}).
 \end{itemize}
 To the best of our knowledge, the minimax rates for nonparametric regression with random design over unknown manifolds have only been worked out for H\"{o}lder classes, and even in this case the calculations are only for $s \leq 2$ bounded derivatives~\citep{bickel2007,yang2016}.
-Our upper bounds confirm that these rates are the same for Sobolev spaces---at least when loss is measured in empirical norm---for the values of $s$ and $m$ mentioned above.
+Our upper bounds confirm that these rates are the same for Sobolev spaces---in estimation, when loss is measured in empirical norm---for the values of $s$ and $m$ mentioned above.
 
@@
 For completeness, we summarize all of our upper bounds---those which match the minimax rates, and those which do not---in Tables~\ref{tbl:estimation_rates} and~\ref{tbl:testing_rates}.
-\begin{table}
+\begin{table}[p]
 	\begin{center}
@@
 \end{table}
-
-\begin{table}
+\begin{table}[p]
 	\begin{center}
@@
 
+\begin{table}[p]
+	\begin{center}
+		\begin{tabular}{p{.14\textwidth} | p{.5\textwidth} }
+		Symbol & Definition
+		\\
+		\hline
+		$\mc{X}$ & domain, either an open set in $\Rd$ (Model~\ref{def:model_flat_euclidean}) or a compact manifold embedded in $\Rd$ (Model~\ref{def:model_manifold}) \\
+		$\nu$ & Lebesgue measure \\
+		$\mu$ & volume form induced by the embedding of $\mc{X}$ into $\Rd$ \\
+		$P$ & probability measure associated with the design points \\
+		$p$ & density of the probability measure, either with respect to $\nu$ (Model~\ref{def:model_flat_euclidean}) or $\mu$ (Model~\ref{def:model_manifold}). \\
+		$L^2(\mc{X})$ & set of square-integrable functions, meaning either $\int_{\mc{X}} f^2 \,d\nu < \infty$ (Model~\ref{def:model_flat_euclidean}) or $\int_{\mc{X}} f^2 \,d\mu < \infty$ (Model~\ref{def:model_manifold}).
+\\
+		$C^k(\mc{X})$ & functions which are $k$-times continuously differentiable in $\mc{X}$ \\
+		$C_c^{\infty}(\mc{X})$ & functions in $C^{\infty}(\mc{X})$ which are compactly supported in $\mc{X}$ \\
+		$H^s(\mc{X})$ & order-s Sobolev space (Definition~\ref{def:sobolev_space} under Model~\ref{def:model_flat_euclidean}, Definition~\ref{def:sobolev_space_manifold} under Model~\ref{def:model_manifold}.) \\
+		$H_0^s(\mc{X})$ & order-s zero-trace Sobolev space (Definition~\ref{def:zero_trace_sobolev_space}) \\
+		$\|\cdot\|_2$ & Euclidean distance \\
+		$d_{\mc{X}}(\cdot,\cdot)$ & geodesic distance \\
+		$B(x,\delta)$ & Ball in Euclidean distance, centered at $x$ with radius $\delta$ \\
+		$B_{\mc{X}}(x,\delta)$ & Ball in geodesic distance
+		\end{tabular}
+	\end{center}
+	\caption{Notation.}
+\end{table}
+
 \paragraph{Technical contributions.}
+\textcolor{red}{(For Siva): This is the section Ryan wanted to cut or put later on. See his Slack message.} \\
 One of our primary technical contributions is a strategy for analyzing regression methods using learned features which leads to sharp rates of convergence.
 
-At an abstract level PCR-LE is a two-stage algorithm: the first stage (LE) uses the observed design points $\{X_1,\ldots,X_n\}$ to learn a feature representation ($X_i \mapsto (v_{1,i},\ldots,v_{K,i})$) which is an empirical approximation to some ideal population-level representation ($X_i \mapsto (\psi_{1}(X_i),\ldots,\psi_{K}(X_i))$);  the second stage then applies a simple method (OLS) to the learned features, and obtains an estimate.
+At an abstract level PCR-LE is a two-stage algorithm.
+The first stage (LE) uses the observed design points $\{X_1,\ldots,X_n\}$ to learn a feature representation ($X_i \mapsto (v_{1,i},\ldots,v_{K,i})$) which is an empirical approximation to some ideal population-level representation ($X_i \mapsto (\psi_{1}(X_i),\ldots,\psi_{K}(X_i))$).
+The second stage then applies a simple method for regression (OLS) using these learned features to obtain an estimate.
 At first blush, one might expect the error for PCR-LE to be likewise decomposed into two parts: first, the error with which the features are learned (i.e.
@@
 \end{equation}
-Thus we obtain an upper bound on the in-sample sample mean squared error $\|\wh{f} - f_0\|_n^2$ that is \emph{independent of the error with which graph Laplacian eigenvectors $v_k$ approximate population-level eigenfunctions $\psi_k$}.
+Thus we obtain an upper bound on the in-sample sample mean squared error $\|\wh{f} - f_0\|_n^2$ that is \emph{independent of the error with which graph Laplacian eigenvectors approximate population-level eigenfunctions}.
 
@@
 The second, arguably more important, is that in order to obtain upper bounds on $\|\wh{f} - f_0\|_n^2$ we do not require that these graph functionals themselves converge to population-level limits, but only that they be stochastically bounded on the right order.
-To derive our ultimate upper bounds on the error of PCR-LE we use some existing results regarding neighborhood graph Laplacians, and prove some new ones which may be of independent interest.
+To derive our ultimate upper bounds on the error of PCR-LE we use some existing results regarding functionals of neighborhood graph Laplacians, and prove some new ones which may be of independent interest.
 
@@
 
+\paragraph{Laplacian smoothing.}
+In a previous paper~\citep{green2021}, we (the authors) considered an alternative method for nonparametric regression via neighborhood graphs: \emph{Laplacian smoothing}, defined as the solution to the following optimization problem,
+\begin{equation}
+\label{eqn:laplacian_smoothing}
+\minimize_{f \in \Reals^n} \|{\bf Y} - f\|_2^2 + \lambda f^{\top} L f.
+\end{equation}
+Laplacian smoothing is penalized method for regression, where the penalty functional $f^{\top} L f$ serves as a discrete approximation to the continuum functional $J(f) := \int \|\nabla f(x)\|^2 p^2(x) \,dx$ \citep{bousquet03}.
+In the univariate setting ($d = 1$), this casts Laplacian smoothing as a discrete and density-weighted alternative to a first-order thin-plate spline estimator, which is defined as the solution to
+\begin{equation}
+\label{eqn:thin_plate_spline}
+\minimize_{f \in H^1(\Reals)} \frac{1}{n}\sum_{i = 1}^{n} (Y_i - f(X_i))^2 + J(f).
+\end{equation}
+When $d = 1$ the first-order thin-plate spline estimator enjoys excellent theoretical properties, such as being minimax optimal over the first-order Sobolev space $H^1(\Reals)$.
+However, when $d \geq 2$ the story changes dramatically: the problem ~\eqref{eqn:thin_plate_spline} is in fact not even well-posed.\footnote{This can be explained by reference to the Sobolev Embedding Theorem, since it is an implication of this theorem that convergence of a sequence of functions $\{f_N\}_{N \in \mathbb{N}} \to f$ in first-order Sobolev norm implies pointwise convergence only when $d = 1$.} In contrast, in this previous paper, we showed that Laplacian smoothing was a well-posed and consistent estimator for all dimensions $d$, and achieved minimax optimal rates for estimation and testing so long as $d \in \{1,2,3,4\}$.
+
+However, Laplacian smoothing neither takes advantage of smooth higher-order derivatives, nor is it provably optimal over $H^1(\mc{X})$ for dimensions $d \geq 5$.
+One of our motivations for considering PCR-LE was to find an estimator which addressed these deficiencies.
+In this work we indeed establish that PCR-LE has much stronger optimality properties than those we derived for Laplacian smoothing, or indeed those known for any other method of regression using neighborhood graphs.
+
+One way to interpret this difference between PCR-LE and Laplacian smoothing is to view the latter as a ridge regression problem.
+This follows from writing the Laplacian smoothing penalty as a (weighted) ridge penalty in the spectral domain,  $f^{\top} L f  = \sum_{k = 1}^{n} \lambda_k(v_k^{\top}f)^2$.
+\citet{dhillon2013} establish conditions under which principal components regression can have smaller risk than ridge regression using the same set of features.
+Viewed in this light, our work shows this phenomenon occurs when the features are eigenvectors of a neighborhood graph Laplacian and the estimand is a function in Sobolev space.
+It also establishes that principal components regression can obtain the minimax rate of convergence even when ridge fails to do so.
+This is not the case if the function class in question is an RKHS~\citep{dicker2017}, and motivates the study of regression over Sobolev spaces in the subcritical regime \textcolor{red}{Siva: is this too fuzzy?}.
+
+\paragraph{Other related work.}
+
 Much of the work regarding regression using neighborhood graph Laplacians deals with \emph{semi-supervised learning}, where in addition to the labeled data $(X_1,Y_1),\ldots,(X_n,Y_n)$ one observes unlabeled points $(X_{n + 1},\ldots,X_{N})$, and the task is to produce an estimate at labeled and unlabeled points alike.
@@
 \label{eqn:graph_laplacian_regularization_ssl}
-\wc{f} := \argmin_{f \in \Reals^N} \|{\bf Y} - f\|_n^2 + \lambda f^{\top} L_N f.
+\minimize_{f \in \Reals^N} \|{\bf Y} - f\|_n^2 + \lambda f^{\top} L_N f.
 \end{equation}
@@
 \label{eqn:laplacian_eigenmaps_ssl}
-\wb{f} := \argmin_{f \in \mathrm{span}\{u_1,\ldots,u_K\}} \|{\bf Y} - f\|_n^2.
+\minimize_{f \in \mathrm{span}\{u_1,\ldots,u_K\}} \|{\bf Y} - f\|_n^2.
 \end{equation}
 \cite{zhou2011,lee2016} analyze SSL-LE in a particular asymptotic regime where the number of labeled points $n$ is held fixed while the number of unlabeled points $N - n \to \infty$.
-They show that the estimator $\wb{f}$ achieves minimax-optimal rates of convergence over $H^s(\mc{X})$ as a function of $n$.
-Of course, as $N - n \to \infty$ the low-frequency eigenvectors of the graph Laplacian $L_N$ all converge to their continuum limits.
-This means in particular that for $K(n) = n^{d/(2s + d)}$,
-\begin{equation*}
-\lim_{N \to \infty} \max_{1 \leq K \leq K(n)} \|\sqrt{N} u_k - \psi_k\|_N^2 = 0,
-\end{equation*}
-and consequently when $\wb{f}$ is optimally tuned, $\lim_{N \to \infty} \|\wb{f} - \wt{f}\|_N^2 = 0$.
-Thus the analysis of SSL-LE reduces to that of population-level spectral series regression.
-The supervised setting (where $N = n$) we consider in this work is very different, and analyzing PCR-LE necessitates an entirely different approach, as explained in the preceding section.
-
-There has been much less work, relatively speaking, regarding regression with neighborhood graph Laplacians in the supervised setting, which is the focus of this work.
-\citet{lee2016} also analyze a variant of PCR-LE, but derive suboptimal rates of convergence.
-\citet{trillos2020} study nonparametric regression via graph Laplacian regularization, where the estimator is computed by solving~\eqref{eqn:graph_laplacian_regularization_ssl}, but with no unlabeled data.
-They establish the uniform upper bound $\max_{i = 1,\ldots,n}|\wc{f}(X_i) - f_0(X_i)| \leq C n^{-2/(2 + d)}$ under the assumption $f_0 \in C^2(\mc{X})$, which is slower than the minimax rate $n^{-4/(4 + d)}$ for this class.
-In a previous paper~\citep{green2021}, we (the authors) also considered nonparametric regression via graph Laplacian regularization, when the regression function $f_0 \in H^1(\mc{X})$ and loss is measured using in-sample mean squared error.
-We found that graph Laplacian regularization is consistent as $n \to \infty$, and optimal (using mean-squared error) over the first-order Sobolev space $H^1(\mc{X})$ when $d \in \{1,2,3,4\}$.
-However, graph Laplacian regularization neither takes advantage of smooth higher-order derivatives, nor is it provably optimal over $H^1(\mc{X})$ for all dimensions $d$.
-One of our motivations for considering PCR-LE was to find an estimator which addressed these deficiencies.
-In this work we indeed find that PCR-LE has much stronger optimality properties than those known for graph Laplacian regularization, or indeed any other method for regression using neighborhood graphs.
+They show that the SSL-LE estimator achieves minimax optimal rates of convergence over $H^s(\mc{X})$ as a function of $n$.
+However, in the particular asymptotic regime when $n$ is fixed and $N - n \to \infty$, the $n$ lowest-frequency eigenvectors of the graph Laplacian $L_N$ all converge to their continuum limits.
+Consequently, the SSL-LE estimator converges to the population-level spectral series estimator, and the analysis of SSL-LE reduces to that of the population-level method.
+As explained in the preceding section, the supervised setting (where $N = n$) we consider in this work is very different, and analyzing PCR-LE necessitates an entirely different approach,
+
+In this supervised setting, there has been relatively little work regarding \emph{random design} regression with neighborhood graph Laplacians .
+Aside from our own work on Laplacian smoothing, summarized above, we highlight two other related papers: \citet{lee2016}, who analyze a variant of PCR-LE, but derive suboptimal rates of convergence, and \citet{trillos2020}, who also study Laplacian smoothing and establish the uniform upper bound $\max_{i = 1,\ldots,n}|\wc{f}(X_i) - f_0(X_i)| \leq C n^{-2/(2 + d)}$ under the assumption $f_0 \in C^2(\mc{X})$, which is slower than the minimax rate $n^{-4/(4 + d)}$ for this class.
 
@@
 \paragraph{Notation.}
-We frequently refer to various classical function classes.
-For an open set $\mc{X}$ equipped with a volume form $d\mu$, we let $L^2(\mc{X})$ denote the set of functions $f$ for which $\|f\|_{L^2(\mc{X})}^2 := \int f^2 \,d\mu  < \infty$, and equip $L^2(\mc{X})$ with the norm $\|\cdot\|_{L^2(\mc{X})}$.
-We define $\dotp{f}{g}_P := \int fg\,dP$, and let $L^2(P)$ contain those functions $f$ for which $\|f\|_P^2 := \dotp{f}{f}_P$ is finite.
-Equivalently, we let $L^2(P_n)$ consist of those ``functions'' $f: \{X_1,\ldots,X_n\} \to \Reals$ for which the empirical norm $\|f\|_{n}^2 := \frac{1}{n}\sum_{i = 1}^{n} \bigl(f(X_i)\bigr)^2 < \infty$.
-When there is no chance of confusion, we will sometimes associate functions in $L^2(P_n)$ with vectors in $\Reals^n$, and vice versa.
+We frequently refer to various classical function classes, starting with the Lebesgue space $L^2(\mc{X})$, defined differently depending on whether $\mc{X} \subseteq \Rd$ is a full-dimensional open set or a compact Riemannian manifold.
+When $\mc{X} \subseteq \Rd$ is a full-dimensional open set, letting $\,d\nu$ denote the Lebesgue measure, the space $L^2(\mc{X})$ refers to the set of $\nu$-measurable functions $f$ for which $\|f\|_{L^2(\mc{X})}^2 := \int f^2 \,d\nu  < \infty$.
+When $\mc{X}$ is a compact Riemannian manifold, letting $\,d\mu$ denote the volume form induced by the embedding of $\mc{X}$ into $\Rd$, the space $L^2(\mc{X})$ refers to the set of $\mu$-measurable functions $f$ for which $\|f\|_{L^2(\mc{X})}^2 := \int f^2 \,d\mu  < \infty$.
+We also define an inner-product over these spaces: for a measure $P$ which admits a density $p$ with respect to $\nu$, we define $\dotp{f}{g}_P := \int f(x)g(x)p(x) \,d\nu(x)$; likewise, if $P$ admits a density $p$ with respect to $\mu$, $\dotp{f}{g}_P := \int f(x)g(x)p(x) \,d\mu(x)$.
+We refer to the norm $\|f\|_P^2 := \dotp{f}{f}_{P}$ as $L^2(P)$-norm.
+
 We use $C^k(\mc{X})$ to refer to functions which are $k$ times continuously differentiable in $\mc{X}$, either for some integer $k \geq 1$ or for $k = \infty$.
@@
 \end{equation} 
-where $V_K \in \Reals^{n \times K}$ is the matrix with $k$th column $V_{K,k} = v_k$. The PCR-LE test statistic is 
+where $V_K \in \Reals^{n \times K}$ is the matrix with $k$th column $V_{K,k} = v_k$.
+The PCR-LE test statistic is defined with respect to the empirical norm $\|f\|_{n}^2 := \frac{1}{n}\sum_{i = 1}^{n} \bigl(f(X_i)\bigr)^2$ of $\wh{f}$:\footnote{Here and throughout, when there is no chance of confusion we will identify vectors $f \in \Reals^n$ with functions $f: \{X_1,\ldots,X_n\} \to \Reals, f(X_i) = f_i$.}
 \begin{equation}
@@
 However, they are unusual in that we treat both the case where $s < d/2$ and thus the Sobolev space $H^s(\mc{X})$ does not continuously embed into $C(\mc{X})$, and the case where $P$ is not the uniform distribution over the unit cube.
-The upper bounds given in this section serve two purposes: first, to clarify what the rates are in these more unusual settings; second, to show that population-level spectral series regression can obtain the optimal rates in such settings.
+The upper bounds given in this section serve two purposes: first, to clarify what the rates are in these less-typically studied settings; second, to show that population-level spectral series regression can obtain the optimal rate.
 This latter point is important since the method we focus on for the most part, PCR-LE, is an empirical approximation to population-level spectral series regression.
@@
 \paragraph{Estimation.}
-Recall the population-level spectral series estimator $\wt{f}$ defined in~\eqref{eqn:population-level_spectral_series}. We now give an upper bound on the $L^2(P)$ risk of $\wt{f}$.
+Recall the population-level spectral series estimator $\wt{f}$ defined in~\eqref{eqn:population-level_spectral_series}.
+We now give an upper bound on the risk of $\wt{f}$, when loss is measured in $L^2(P)$ norm.
 \begin{proposition}
@@
 Nevertheless, for random design regression with error measured in squared $L^2(P)$-norm, the population-level spectral series estimator $\wt{f}$ obtains the standard minimax rates $n^{-2s/(2s + d)}$ for all values of $s$ and $d$.
-Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and \ref{thm:laplacian_eigenmaps_estimation_ho} show that the same is true with respect to PCR-LE, when error is measured in squared $L^2(P_n)$-norm.
+Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and \ref{thm:laplacian_eigenmaps_estimation_ho} show that the same is true with respect to PCR-LE, when error is measured in empirical norm.
 	\item \emph{Smoothness of design density}.
@@
 \end{enumerate}
-It is worth pointing out that we do not need to establish the pointwise estimate $L_{P,\varepsilon}^sf \to \sigma_{\eta}^s\Delta_{P}^sf$ in $L^2(P)$.
+It is worth pointing out that we do not need to establish the pointwise estimate $L_{P,\varepsilon}^sf \to \sigma_{\eta}^s\Delta_{P}^sf$ in $L^2(P)$ norm.
 If we had such an estimate, it would immediately follow that $\dotp{L_{P,\varepsilon}^sf}{f}_{P} \to \sigma_{\eta}^s\dotp{\Delta_P^sf}{f}_{P}$.
@@
 \paragraph{Tuning parameters.}
-Our first two experiments demonstrate that PCR-LE methods have comparable statistical performance to population-leve spectral series methods.
+Our first two experiments demonstrate that PCR-LE methods have comparable statistical performance to population-level spectral series methods.
 PCR-LE depends on two tuning parameters, and in our final experiment we investigate the importance of both, focusing now on estimation.
@@
 Of course, there do exist other methods for nonparametric regression which achieve optimal rates of convergence under similar (or indeed weaker) conditions on $p$.
-These include other graph-based approaches---graph Laplacian regularization---methods besides spectral series methods---e.g.
+These include other graph-based approaches---Laplacian smoothing---methods besides spectral series methods---e.g.
 kernel smoothing, local polynomial regression, thin-plate splines---and continuum spectral projection methods which use the eigenfunctions of an operator defined independently of $p$.
@@
 	\item \emph{Optimality over high-dimensional Sobolev spaces}.
-As mentioned in the introduction, graph Laplacian regularization provably achieves minimax optimal rates over $H^1(\mc{X})$ only when $d \in \{1,2,3,4\}$ \citep{sadhanala16, green2021}; when $d \geq 5$, graph Laplacian regularization appears to be suboptimal.
-In contrast, PCR-LE is optimal over $H^1(\mc{X})$ for all dimensions $d$.
-An even more stark contrast can be drawn between PCR-LE (and other spectral series estimators) and the first order thin-plate spline, the solution of the variational problem
-	\begin{equation}
-	\underset{f \in H^1(\mc{X})}{\textrm{minimize}} \quad  \|{\bf Y} - f\|_n^2 + \lambda \|f\|_{H^1(\Reals^d)}^2,
-	\end{equation}
-	which is optimal and indeed well-posed only when $d = 1$.
-This appears to be related to a general phenomenon described by~\cite{dhillon2013,dicker2017}, in which principal components analysis followed by ordinary least-squares is shown to be always competitive with, and sometimes much better than, ridge regression.
+As mentioned in the introduction, Laplacian smoothing (defined via~\eqref{eqn:laplacian_smoothing}) provably achieves minimax optimal rates over $H^1(\mc{X})$ only when $d \in \{1,2,3,4\}$ \citep{sadhanala16, green2021}.
+In contrast, PCR-LE is optimal over $H^1(\mc{X})$ for all dimensions $d$, and also over the higher-order Sobolev spaces $H^s(\mc{X})$.
 	\item \emph{Manifold adaptivity}.

### 2021-10-29 manuscripts/laplacian_eigenmaps/appendix.tex
Responded to all of Siva's comments.

@@
 \noindent
+
+\section{Notation Table}
+\label{sec:notation_table}
+
+\begin{table}[h]
+	\begin{center}
+		\begin{tabular}{p{.14\textwidth} | p{.5\textwidth} }
+			Symbol & Definition
+			\\
+			\hline
+			$\mc{X}$ & domain, either an open set in $\Rd$ (Model~\ref{def:model_flat_euclidean}) or a compact manifold embedded in $\Rd$ (Model~\ref{def:model_manifold}) \\
+			$\nu$ & Lebesgue measure \\
+			$\mu$ & volume form induced by the embedding of $\mc{X}$ into $\Rd$ \\
+			$P$ & probability measure associated with the design points \\
+			$p$ & density of the probability measure, either with respect to $\nu$ (Model~\ref{def:model_flat_euclidean}) or $\mu$ (Model~\ref{def:model_manifold}). \\
+			$L^2(\mc{X})$ & set of square-integrable functions, meaning either $\int_{\mc{X}} f^2 \,d\nu < \infty$ (Model~\ref{def:model_flat_euclidean}) or $\int_{\mc{X}} f^2 \,d\mu < \infty$ (Model~\ref{def:model_manifold}).
+\\
+			$C^k(\mc{X})$ & functions which are $k$-times continuously differentiable in $\mc{X}$ \\
+			$C_c^{\infty}(\mc{X})$ & functions in $C^{\infty}(\mc{X})$ which are compactly supported in $\mc{X}$ \\
+			$H^s(\mc{X})$ & order-s Sobolev space (Definition~\ref{def:sobolev_space} under Model~\ref{def:model_flat_euclidean}, Definition~\ref{def:sobolev_space_manifold} under Model~\ref{def:model_manifold}.) \\
+			$H_0^s(\mc{X})$ & order-s zero-trace Sobolev space (Definition~\ref{def:zero_trace_sobolev_space}) \\
+			$\|\cdot\|_2$ & Euclidean distance \\
+			$d_{\mc{X}}(\cdot,\cdot)$ & geodesic distance \\
+			$B(x,\delta)$ & Ball in Euclidean distance, centered at $x$ with radius $\delta$ \\
+			$B_{\mc{X}}(x,\delta)$ & Ball in geodesic distance
+		\end{tabular}
+	\end{center}
+	\caption{Notation.}
+\end{table}
 

### 2021-10-29 manuscripts/laplacian_eigenmaps/arxiv/laplacian_eigenmaps_regression.tex
Responded to all of Siva's comments.

@@
 PCR-LE works by projecting a vector of observed responses ${\bf Y} = (Y_1,\ldots,Y_n)$ onto a subspace spanned by certain eigenvectors of a neighborhood graph Laplacian.
-We show that PCR-LE achieves minimax rates of convergence over continuum function classes.
-Specifically, we consider the problem of random design regression over Sobolev spaces $H^s(\mc{X})$---where $\mc{X} \subseteq \Rd$ is the support of the design distribution $P$---and show that under sufficient smoothness conditions on the design density $p$, PCR-LE achieves the optimal rates for both estimation (where the optimal rate in squared $L^2$ norm is known to be $n^{-2s/(2s + d)}$) and goodness-of-fit testing ($n^{-4s/(4s + d)}$).
+We show that PCR-LE achieves minimax rates of convergence for random design regression over Sobolev spaces.
+Under sufficient smoothness conditions on the design density $p$, PCR-LE achieves the optimal rates for both estimation (where the optimal rate in squared $L^2$ norm is known to be $n^{-2s/(2s + d)}$) and goodness-of-fit testing ($n^{-4s/(4s + d)}$).
 We also show that PCR-LE is \emph{manifold adaptive}: that is, we consider the situation where $\mc{X}$ is a manifold of small intrinsic dimension $m$, and give upper bounds establishing that PCR-LE achieves the faster minimax estimation ($n^{-2s/(2s + m)}$) and testing ($n^{-4s/(4s + m)}$) rates of convergence.
-Interestingly, these rates are almost always much faster than the known rates of convergence of graph Laplacian eigenvectors to their population-level limits; in other words, for this problem regression with learned features appears to be much easier, statistically speaking, than feature learning itself.
+Interestingly, these rates are almost always much faster than the known rates of convergence of graph Laplacian eigenvectors to their population-level limits; in other words, for this problem regression with estimated features appears to be much easier, statistically speaking, than estimating the features itself.
 We support these theoretical results with empirical evidence.
@@
 
+\newpage
+
 % Appendix

### 2021-10-29 manuscripts/laplacian_eigenmaps/main_text.tex
Responded to all of Siva's comments.

@@
 	\item First, LE forms a \emph{neighborhood graph} $G = (V,W)$ over the points $\{X_1,\ldots,X_n\}$.
-The graph $G$ is an undirected, weighted graph, with vertices $V = \{X_1,\ldots,X_n\}$, and weighted edges $W_{ij}$ which correspond to the proximity between points $X_i$ and $X_j$, say in Euclidean distance.
+The graph $G$ is an undirected, weighted graph, with vertices $V = \{X_1,\ldots,X_n\}$, and weighted edges $W_{ij}$ which correspond to the proximity between points $X_i$ and $X_j$.
 	\item Next, LE forms an (unweighted) \emph{graph Laplacian} matrix $L \in \Reals^{n \times n}$, a symmetric and diagonally dominant matrix with diagonal elements $L_{ii} = \sum_{j = 1}^{n} W_{ij}$, and off-diagonal elements $L_{ij} = -W_{ij}$.
@@
 \end{enumerate} 
-A natural way to use LE is by taking the vectors $\{(v_{1,i},\ldots,v_{K,i})\}_{i = 1}^{n}$ to be features in a downstream supervised learning algorithm.
-In this paper, we study a simple method along these lines: Principal Components Regression with Laplacian-Eigenmaps (PCR-LE), a method for nonparametric regression which operates by running ordinary least squares (OLS) using the features learned by LE.
+A natural way to use LE is by taking the vectors $\{(v_{1,i},\ldots,v_{K,i})\}_{i = 1}^{n}$ to be features in a downstream regression algorithm.
+In this paper, we study a simple method along these lines: Principal Components Regression with Laplacian-Eigenmaps (PCR-LE), a method for nonparametric regression which operates by running ordinary least squares (OLS) using the features output by LE.
 Given pairs of design points and responses $(X_1,Y_1),\ldots, (X_n,Y_n)$, PCR-LE computes an estimate $\wh{f} \in \Reals^n$,
@@
 \end{equation}  
-(Here $\mathrm{div}$ stands for the divergence operator, and $\nabla$ for the gradient.) These eigenfunctions in turn characterize various interesting structural aspects of $p$, such as the location and number of high- and low-density regions, the shape and intrinsic dimension of its support, and so forth.
-
-These aforementioned works explain what kind of representation LE learns, and the accuracy with which it learns this representation.
+(Here $\mathrm{div}$ stands for the divergence operator, and $\nabla$ for the gradient.
+See~\eqref{eqn:laplace_beltrami_eigenproblem} for the formal definition of eigenpairs $(\rho_k,\psi_k)$.) These eigenfunctions in turn characterize various interesting structural aspects of $p$, such as the location and number of high- and low-density regions, the shape and intrinsic dimension of its support, and so forth.
+
+These aforementioned works justify LE as method for data representation, by establishing that each feature vector $(v_{1,i},\ldots,v_{K,i})$ serves an empirical approximation to an idealized representation $(\psi_1(X_i),\ldots,\psi_K(X_i))$.
+They also provide quantitative guarantees for the accuracy with which LE approximates this ideal representation.
 However, this theory does not focus on the statistical properties of PCR-LE for classical regression problems such as estimation and testing.
@@
 
-At first glance, the optimality of PCR-LE is surprising.
+% (AG 10/28/21): I comment out the paragraph below. I decided the better approach was to first give our results, then comment on their perspective. 
+% At first glance, the optimality of PCR-LE is surprising.
 As a function of the sample size $n$, known rates of convergence of LE are much slower than the minimax optimal rates of convergence over $H^s(\mc{X})$ (roughly speaking $n^{-2s/(2s + d)}$ for estimation and $n^{-4s/(4s + d)}$ for testing).
@@
 We study regression over Sobolev spaces in part because, generally speaking, the minimax rates are well-understood; as mentioned before, when the domain $\mc{X}$ is full-dimensional they are $n^{-2s/(2s + d)}$ for estimation, and $n^{-4s/(4s + d)}$ for testing.
-For this reason, regression over Sobolev spaces is a good setting in which to see whether PCR-LE measures up to more standard minimax optimal approaches, which have strong theoretical guarantees but are less often used in practice because their applicability is more limited.
-\textcolor{red}{(TODO): This last clause still seems weak to me.
-Does Siva agree?}
-
-Moreover, we view PCR-LE as being particularly well-suited for regression over Sobolev spaces due to their close connection with \emph{spectral series regression}.
-Spectral series regression computes empirical Fourier coefficients and truncates to the $K$-lowest frequency eigenfunctions, producing the estimate
+For this reason, regression over Sobolev spaces is a good setting in which to see whether PCR-LE measures up to more standard minimax optimal approaches, which have strong theoretical guarantees but are less often used in practice.
+We give a more specific comparison between PCR-LE and some of these more classical methods in Section~\ref{sec:discussion}.
+
+We also view PCR-LE as being particularly well-suited for regression over Sobolev spaces due to their close connection with \emph{spectral series regression}.
+Spectral series regression computes generalized empirical Fourier coefficients $\wt{a}_k := \frac{1}{n}\sum_{i = 1}^{n} Y_i \psi_k(X_i)$, and truncates to the $K$-lowest frequency eigenfunctions of $\Delta_P$, producing the estimate
 \begin{equation}
 \label{eqn:population-level_spectral_series}
-\wt{a}_k = \frac{1}{n}\sum_{i = 1}^{n} Y_i \psi_k(X_i), \quad \wt{f}(x) = \sum_{k = 1}^{K} \wt{a}_k \psi_k(x).
-\end{equation} 
-Under appropriate boundary conditions, the Sobolev spaces $H^s(\mc{X})$ consists of function $f = \sum_{k} a_k \psi_k \in L^2(\mc{X})$ with generalized Fourier coefficients $\{a_k\}_{k = 1}^{\infty}$ that satisfy the decay condition $\sum_{k} a_k \rho_k^2 < \infty$ (See Section~\ref{subsec:spectral_projection} for more details).
+\wt{f}(x) = \sum_{k = 1}^{K} \wt{a}_k \psi_k(x).
+\end{equation} 
+Spectral series regression is intrinsically linked with Sobolev spaces.
+That is because under appropriate boundary conditions, a unit ball in the order-$s$ Sobolev space consists of functions $f = \sum_{k} a_k \psi_k \in L^2(\mc{X})$ for which the generalized Fourier coefficients $\{a_k\}_{k = 1}^{\infty}$ satisfy the decay condition $\sum_{k} a_k^2 \rho_k^s \leq C$ (See Section~\ref{subsec:spectral_projection} for more details).
 This decay condition justifies the truncated series estimator \eqref{eqn:population-level_spectral_series}, since it means the truncation will incur only a limited amount of bias for any $f_0 \in H^s(\mc{X})$.
-For this reason spectral series regression over Sobolev spaces has been well-studied since at least \citet{rice1984},\footnote{And proposed much earlier in the context of density estimation by \citet{cencov1962}.} and its optimality properties are by this point well-understood.
-
-PCR-LE serves as an empirical approximation to spectral series regression, since the eigenvectors $v_k$ are empirical approximations to the eigenfunctions $\psi_k$.
+For this reason spectral series regression over Sobolev spaces has been well-studied---at least when $\mc{X} = [0,1]^d$---since at least \citet{rice1984},\footnote{And proposed much earlier in the context of density estimation by \citet{cencov1962}.} and its optimality properties are by this point generally well-understood.
+
+PCR-LE serves as an empirical approximation to spectral series regression, since as already mentioned the eigenvectors $v_k$ are empirical approximations to the eigenfunctions $\psi_k$ of $\Delta_P$.
 Viewed in this light, a major advantage of PCR-LE is that it operates without needing knowledge of the design distribution $P$.
-In contrast, the spectral series estimator defined in~\eqref{eqn:population-level_spectral_series} relies on diagonalizing the density-weighted Laplacian $\Delta_P$.
-In our context, where $P$ is treated as unknown,~\eqref{eqn:population-level_spectral_series} must be viewed as an oracle method, and to emphasize this we henceforth refer to it as \emph{population-level spectral series regression}.
+This is an advantage because in our context $P$ is treated as an unknown and potentially complex distribution, and for example can be highly non-uniform, have a complicated support which may be a submanifold of $\Rd$, or both.
+In contrast, spectral series regression relies on diagonalizing the density-weighted Laplacian $\Delta_P$, and in our context must be viewed as an oracle method; to emphasize this we henceforth refer to the estimator defined in~\eqref{eqn:population-level_spectral_series} as \emph{population-level spectral series regression}.
 On the other hand, intuitively PCR-LE incurs some extra error by using an empirical approximation to the underlying basis $\{\psi_k\}_{k = 1}^{\infty}$: our work shows that in many cases, this extra error is not enough to change the overall rate of convergence.
@@
 For completeness, we summarize all of our upper bounds---those which match the minimax rates, and those which do not---in Tables~\ref{tbl:estimation_rates} and~\ref{tbl:testing_rates}.
-\begin{table}[p]
+
+\begin{table}
 	\begin{center}
@@
 \end{table}
-\begin{table}[p]
+\begin{table}
 	\begin{center}
@@
 
-\begin{table}[p]
-	\begin{center}
-		\begin{tabular}{p{.14\textwidth} | p{.5\textwidth} }
-		Symbol & Definition
-		\\
-		\hline
-		$\mc{X}$ & domain, either an open set in $\Rd$ (Model~\ref{def:model_flat_euclidean}) or a compact manifold embedded in $\Rd$ (Model~\ref{def:model_manifold}) \\
-		$\nu$ & Lebesgue measure \\
-		$\mu$ & volume form induced by the embedding of $\mc{X}$ into $\Rd$ \\
-		$P$ & probability measure associated with the design points \\
-		$p$ & density of the probability measure, either with respect to $\nu$ (Model~\ref{def:model_flat_euclidean}) or $\mu$ (Model~\ref{def:model_manifold}). \\
-		$L^2(\mc{X})$ & set of square-integrable functions, meaning either $\int_{\mc{X}} f^2 \,d\nu < \infty$ (Model~\ref{def:model_flat_euclidean}) or $\int_{\mc{X}} f^2 \,d\mu < \infty$ (Model~\ref{def:model_manifold}).
-\\
-		$C^k(\mc{X})$ & functions which are $k$-times continuously differentiable in $\mc{X}$ \\
-		$C_c^{\infty}(\mc{X})$ & functions in $C^{\infty}(\mc{X})$ which are compactly supported in $\mc{X}$ \\
-		$H^s(\mc{X})$ & order-s Sobolev space (Definition~\ref{def:sobolev_space} under Model~\ref{def:model_flat_euclidean}, Definition~\ref{def:sobolev_space_manifold} under Model~\ref{def:model_manifold}.) \\
-		$H_0^s(\mc{X})$ & order-s zero-trace Sobolev space (Definition~\ref{def:zero_trace_sobolev_space}) \\
-		$\|\cdot\|_2$ & Euclidean distance \\
-		$d_{\mc{X}}(\cdot,\cdot)$ & geodesic distance \\
-		$B(x,\delta)$ & Ball in Euclidean distance, centered at $x$ with radius $\delta$ \\
-		$B_{\mc{X}}(x,\delta)$ & Ball in geodesic distance
-		\end{tabular}
-	\end{center}
-	\caption{Notation.}
-\end{table}
-
-\paragraph{Technical contributions.}
-\textcolor{red}{(For Siva): This is the section Ryan wanted to cut or put later on. See his Slack message.} \\
-One of our primary technical contributions is a strategy for analyzing regression methods using learned features which leads to sharp rates of convergence.
-
-At an abstract level PCR-LE is a two-stage algorithm.
-The first stage (LE) uses the observed design points $\{X_1,\ldots,X_n\}$ to learn a feature representation ($X_i \mapsto (v_{1,i},\ldots,v_{K,i})$) which is an empirical approximation to some ideal population-level representation ($X_i \mapsto (\psi_{1}(X_i),\ldots,\psi_{K}(X_i))$).
-The second stage then applies a simple method for regression (OLS) using these learned features to obtain an estimate.
-At first blush, one might expect the error for PCR-LE to be likewise decomposed into two parts: first, the error with which the features are learned (i.e.
-error due to sampling variation in $X$), second, the error with which, given ideal population-level features, the regression function is learned (i.e.
-error due to noise in $Y$.)
+\paragraph{Perspective: Regression with Empirical Features vs. Feature Estimation.}
+We now pause for a moment, to emphasize that in a certain respect the aforementioned rates of convergence for PCR-LE are quite surprising.
+Remember that PCR-LE is a regression method using features (eigenvectors $v_k$ of the graph Laplacian $L$) which are themselves empirical estimates of population-level quantities (eigenfunctions $\psi_k$ of the density-weighted Laplacian $\Delta_P$).
+It seems reasonable to expect that the error of PCR-LE should be decomposed into two parts: first, the error with which these empirically-derived features estimate their continuum limits; second, the error with which, given ideal population-level features, the regression function is learned.
 
 Crucially, our analysis \emph{does not} work in this way.
-This is important because, as already mentioned, all known upper bounds on the error with which the learned features (Laplacian eigenvectors) approximate their population-level limits (continuum Laplacian eigenfunctions) are much larger than the minimax rates of convergence over $H^s(\mc{X})$.
-For example,
-\begin{itemize}
-	\item The best known rates of convergence for graph Laplacian eigenvectors (due to~\cite{cheng2021}) are
-	\begin{equation}
-	\label{eqn:eigenvector_convergence}
-	\max_{1 \leq k \leq K}\|\sqrt{n}v_k - \psi_k\|_n^2 \leq C_{K} n^{-2/(4 + d)},
-	\end{equation}
-	where the constant $C_K$ depends on $K$ in an unknown manner.
-	\item The dependence of $C_K$ on $K$ is relevant for our interests, since in order to optimally trade off bias and variance we will need to choose $K$ is growing with $n$.
-The earlier works of \cite{burago2014,trillos2019} give upper bounds which, although implying weaker rates of convergence as a function of $n$, do keep track of $C_K$, and suggest that it should depend on the inverse of the spectral gap
-	\begin{equation}
-	\label{eqn:spectral_gap}
-	C_K = \frac{C}{\rho_{K + 1} - \rho_K} \geq C K^{1 - 2/d},
-	\end{equation}
-	where $C$ is a constant depending only on the dimension $d$, and the latter inequality follows whenever the eigenvalues $\rho_K$ obey the Weyl's Law asymptotic scaling $\rho_K \asymp K^{2/d}$.
-Thus, for a given $K(n)$ growing with $n$, when $d \geq 3$, the rates of convergence of $v_{K(n)} - \psi_{K(n)}$ are even slower than $n^{-2/(4 + d)}$.
-\end{itemize}
-Although these upper bounds may not reflect the true rate of convergence of graph Laplacian eigenvectors---this is still an active area of research, and no lower bounds are known---it seems very unlikely that the true rate matches the minimax estimation rate $n^{-2s/(2s + d)}$, which after all approaches the dimension-free rate $1/n$ for large values of $s$.
-
-Instead of relying on convergence of eigenvectors to eigenfunctions, our analysis proceeds via a bias-variance decomposition at the level of the graph. Focusing for simplicity on estimation, this is 
-\begin{equation}
-\label{eqn:bias_variance_estimation}
-\|f_0 - \wh{f}\|_n^2 \leq 2\bigl(\|f_0 - V_KV_K^{\top}f_0\|_n^2 + \|V_KV_K^{\top}({\bf Y} - f_0)\|_n^2\bigr) 
-\end{equation}
-The second term in~\eqref{eqn:bias_variance_estimation} is the variance, and as usual for OLS estimates depends only on the degrees of freedom $\mathrm{df}(\wh{f}) = \mathrm{tr}(V_KV_K^{\top}) = K$.
-More surprisingly, the first term (squared-bias) can also be upper bounded without appealing to the eigenfunctions $\psi_k$.
-Letting $S = L^{\dagger}$ be the pseudo-inverse of the graph Laplacian, a few lines of algebra give
-\begin{equation}
-\label{eqn:bias_estimation}
-\begin{aligned}
-\|(I - V_KV_K^{\top})f_0\|_n^2 & = \|(I - V_KV_K^{\top})S^{s/2}L^{s/2}f_0\|_n^2 \leq \|(I - V_KV_K^{\top})S^{s/2}\|_{op}^2 \|L^{s/2}f_0\|_n^2 = \frac{f_0^{\top} L^s f_0}{n \lambda_{K + 1}^{s}}.
-\end{aligned}
-\end{equation}
-Thus we obtain an upper bound on the in-sample sample mean squared error $\|\wh{f} - f_0\|_n^2$ that is \emph{independent of the error with which graph Laplacian eigenvectors approximate population-level eigenfunctions}.
-
-Instead, our upper bound on the error of PCR-LE is determined by a pair of graph functionals: the quadratic form $f_0^{\top}L^s f_0$, and the graph Laplacian eigenvalue $\lambda_{K + 1}$.
-Theoretically speaking, this brings two advantages over directly analyzing convergence of eigenvectors to eigenfunctions.
-The first is that sharper rates  of convergence are known for these graph functionals than for the eigenvectors $v_k$.
-The second, arguably more important, is that in order to obtain upper bounds on $\|\wh{f} - f_0\|_n^2$ we do not require that these graph functionals themselves converge to population-level limits, but only that they be stochastically bounded on the right order.
-To derive our ultimate upper bounds on the error of PCR-LE we use some existing results regarding functionals of neighborhood graph Laplacians, and prove some new ones which may be of independent interest.
-
-A last observation regarding the difference between feature learning and regression using learned features.
-To balance the bias and variance terms in~\eqref{eqn:bias_variance_estimation}, we will end up using $K(n) := n^{d/(2s + d)}$ eigenvectors.
-Although this is the usual choice for spectral series estimation over order-$s$ Sobolev spaces, in the context of PCR-LE it leads to a striking conclusion.
-Namely, examining~\eqref{eqn:eigenvector_convergence} and~\eqref{eqn:spectral_gap}, we see that for all dimensions $d \geq 3$, known upper bounds on $\|\sqrt{n} v_{K(n)} - \psi_{K(n)}\|_n^2$ do not converge to $0$ as $n \to \infty$.
-In words, it is not merely the case that PCR-LE converges at a faster rate than LE itself, but that PCR-LE can profitably use eigenvectors for regression which may be inconsistent estimates of their population-level limits.
-
-To summarize, our work demonstrates, broadly speaking, that regression using learned features can be analyzed independent of the accuracy of the features themselves.
-Regression using learned features is a general and widely applied paradigm, and we believe this observation may have consequences outside of its application to PCR-LE in this work.
+This is important because all known upper bounds on the rates at which $v_k \to \psi_K$ as $n \to \infty$ are much slower than the minimax rates for regression over Sobolev classes.
+For instance, the best currently known upper bound on the empirical $L^2$ error $\frac{1}{n}\sum_{i = 1}^{n}(\sqrt{n} v_{k,i} - \psi_k(X_i))^2$ is only on the order of $n^{-2/(4 + d)}$~\cite{cheng2021}, which is slower than the minimax estimation rate over $H^s(\mc{X})$ for any $s \in \mathbb{N}, s \geq 1$.\footnote{To make matters worse, PCR-LE, when deployed optimally, does not use a single eigenvector $v_k$ for a fixed index $k \in \mathbb{N}$, but rather many eigenvectors $v_1,\ldots,v_K$ with $K$ growing in $n$.
+As $K$ grows larger, the rate at which $v_K \to \psi_K$ gets slower, since the population-level object being estimated is less regular; see~\citep{burago2014,trillos2019}.} Although this upper bound may not reflect the true rate of convergence of graph Laplacian eigenvectors---this is still an active area of research, and no lower bounds are known---it seems very unlikely that the true rate matches the minimax estimation rate $n^{-2s/(2s + d)}$, which after all approaches the dimension-free rate $1/n$ for large values of $s$.
+The bottom line is that the rate at which graph Laplacian eigenvectors are known to converge to density-weighted Laplacian eigenfunctions is too slow to explain the upper bounds we establish for PCR-LE.
+
+Instead of relying on convergence of eigenvectors to eigenfunctions, our analysis proceeds via a bias-variance decomposition at the level of the graph.
+As usual for OLS estimates, the variance term depends only on the degrees of freedom $\mathrm{df}(\wh{f}) = \mathrm{tr}(V_KV_K^{\top}) = K$.
+More surprisingly, the bias can also be upper bounded without appealing to concentration of eigenvectors $v_1,\ldots,v_K$ around eigenfunctions $\psi_1,\ldots,\psi_K$; for instance, we show in Lemma~\ref{lem:fixed_graph_estimation} that for estimation the squared bias is at most on the order of $f_0^{\top} L^s f_0/(n\lambda_{K + 1}^s)$.
+
+Ultimately our upper bound on the error of PCR-LE is determined entirely by a pair of graph functionals: the quadratic form $f_0^{\top}L^s f_0$, and the graph Laplacian eigenvalue $\lambda_{K + 1}$.
+This brings a couple of advantages.
+First, it eliminates the need to analyze convergence of eigenvectors to eigenfunctions, which is critical in order to get sufficiently fast rates of convergence for PCR-LE, as we have already explained.
+Instead, we only have to consider these two graph functionals, both of which are known to converge at faster rates than graph Laplacian eigenvectors.
+Second, in order to obtain upper bounds on $\|\wh{f} - f_0\|_n^2$ we do not require that these graph functionals themselves converge to population-level limits, but only that they be stochastically bounded on the right order, which is a much weaker requirement.
+To derive our upper bounds on the error of PCR-LE, we directly analyze the quadratic form $f_0^{\top}L^s f_0$ and the eigenvalue $\lambda_{K + 1}$, using some existing results as well as deriving some new ones which may be of independent interest.
+
+To summarize, our work demonstrates, broadly speaking, that regression using estimated features can be analyzed independently from the estimation error of the features themselves.
+Regression using learned features---that is, a feature representation derived from the data itself---is a general and widely applied paradigm, and we believe this observation may have consequences outside of its application to PCR-LE in this work.
 
@@
 When $d = 1$ the first-order thin-plate spline estimator enjoys excellent theoretical properties, such as being minimax optimal over the first-order Sobolev space $H^1(\Reals)$.
-However, when $d \geq 2$ the story changes dramatically: the problem ~\eqref{eqn:thin_plate_spline} is in fact not even well-posed.\footnote{This can be explained by reference to the Sobolev Embedding Theorem, since it is an implication of this theorem that convergence of a sequence of functions $\{f_N\}_{N \in \mathbb{N}} \to f$ in first-order Sobolev norm implies pointwise convergence only when $d = 1$.} In contrast, in this previous paper, we showed that Laplacian smoothing was a well-posed and consistent estimator for all dimensions $d$, and achieved minimax optimal rates for estimation and testing so long as $d \in \{1,2,3,4\}$.
+However, when $d \geq 2$ the story changes dramatically: the problem ~\eqref{eqn:thin_plate_spline} is in fact not even well-posed.\footnote{This can be explained by reference to the Sobolev Embedding Theorem, since it is an implication of this theorem that convergence of a sequence of functions $\{f_N\}_{N \in \mathbb{N}} \to f$ in first-order Sobolev norm implies pointwise convergence only when $d = 1$.} In contrast, in this previous paper, we showed that Laplacian smoothing was a well-posed and consistent estimator for any (fixed) dimension $d$, and achieved minimax optimal rates for estimation and testing so long as $d \in \{1,2,3,4\}$.
 
@@
 It also establishes that principal components regression can obtain the minimax rate of convergence even when ridge fails to do so.
-This is not the case if the function class in question is an RKHS~\citep{dicker2017}, and motivates the study of regression over Sobolev spaces in the subcritical regime \textcolor{red}{Siva: is this too fuzzy?}.
+Interestingly, this is not the case if the function class in question is an RKHS~\citep{dicker2017}, and further motivates the study of regression over Sobolev spaces in the subcritical regime, where surprising new phenomena emerge.
 
@@
 \label{eqn:graph_laplacian_regularization_ssl}
-\minimize_{f \in \Reals^N} \|{\bf Y} - f\|_n^2 + \lambda f^{\top} L_N f.
+\minimize_{f \in \Reals^N} \sum_{i = 1}^{n}(Y_i - f_i)^2 + \lambda f^{\top} L_N f.
 \end{equation}
@@
 One solution to this problem is to instead use Laplacian Eigenmaps for semi-supervised learning (SSL-LE), i.e.
-compute the eigendecomposition $L_N = \sum_{k = 1}^{N} \lambda_k u_k u_k^{\top}$ and solve the problem
+compute the eigendecomposition $L_N = \sum_{k = 1}^{N} \lambda_k u_k u_k^{\top}$ and, letting $U \in \Reals^{n \times K}$ be the matrix with entries $U_{ik} = u_{k,i}$ and columns $U_1,\ldots,U_K$, solve the problem
 \begin{equation}
 \label{eqn:laplacian_eigenmaps_ssl}
-\minimize_{f \in \mathrm{span}\{u_1,\ldots,u_K\}} \|{\bf Y} - f\|_n^2.
+\minimize_{f \in \mathrm{span}\{U_1,\ldots,U_K\}} \sum_{i = 1}^{n}(Y_i - f_i)^2.
 \end{equation}
 \cite{zhou2011,lee2016} analyze SSL-LE in a particular asymptotic regime where the number of labeled points $n$ is held fixed while the number of unlabeled points $N - n \to \infty$.
-They show that the SSL-LE estimator achieves minimax optimal rates of convergence over $H^s(\mc{X})$ as a function of $n$.
+They show that the SSL-LE estimator achieves minimax optimal rates---as a function of the number of labeled points $n$---over Sobolev spaces.
 However, in the particular asymptotic regime when $n$ is fixed and $N - n \to \infty$, the $n$ lowest-frequency eigenvectors of the graph Laplacian $L_N$ all converge to their continuum limits.
 Consequently, the SSL-LE estimator converges to the population-level spectral series estimator, and the analysis of SSL-LE reduces to that of the population-level method.
-As explained in the preceding section, the supervised setting (where $N = n$) we consider in this work is very different, and analyzing PCR-LE necessitates an entirely different approach,
+As we have already explained, the supervised setting (where $N = n$) we consider in this work is very different, and analyzing PCR-LE necessitates an entirely different approach,
 
 In this supervised setting, there has been relatively little work regarding \emph{random design} regression with neighborhood graph Laplacians .
-Aside from our own work on Laplacian smoothing, summarized above, we highlight two other related papers: \citet{lee2016}, who analyze a variant of PCR-LE, but derive suboptimal rates of convergence, and \citet{trillos2020}, who also study Laplacian smoothing and establish the uniform upper bound $\max_{i = 1,\ldots,n}|\wc{f}(X_i) - f_0(X_i)| \leq C n^{-2/(2 + d)}$ under the assumption $f_0 \in C^2(\mc{X})$, which is slower than the minimax rate $n^{-4/(4 + d)}$ for this class.
+Aside from our own work on Laplacian smoothing, summarized above, we highlight two other related papers: \citet{lee2016}, who analyze a variant of PCR-LE, but derive suboptimal rates of convergence, and \citet{trillos2020}, who study Laplacian smoothing and establish the uniform upper bound $\max_{i = 1,\ldots,n}|\wc{f}(X_i) - f_0(X_i)| \leq C n^{-2/(2 + d)}$ under the assumption $f_0 \in C^2(\mc{X})$, which is slower than the minimax rate $n^{-4/(4 + d)}$ for this class.
 
@@
 \paragraph{Notation.}
+We now introduce some notation; for ease of reference, we include a table summarizing notation in Appendix~\ref{sec:notation_table}.
+
 We frequently refer to various classical function classes, starting with the Lebesgue space $L^2(\mc{X})$, defined differently depending on whether $\mc{X} \subseteq \Rd$ is a full-dimensional open set or a compact Riemannian manifold.
@@
 
-\paragraph{Spectrally defined Sobolev spaces.}
+\subsubsection{Spectrally defined Sobolev spaces}
 Let $\mc{X}$ be an open domain which satisfies the conditions of Model~\ref{def:model_flat_euclidean}.
@@
 \end{equation}
-Under Model~\ref{def:model_flat_euclidean}, the eigenvector equation~\eqref{eqn:laplace_beltrami_eigenproblem} has enumerable solutions $(\rho_1,\psi_1),(\rho_2,\psi_2),\ldots$, sorted as usual in ascending order of eigenvalue~\citep{garciatrillos18,trillos2019}.
+Under Model~\ref{def:model_flat_euclidean}, the eigenvector equation~\eqref{eqn:laplace_beltrami_eigenproblem} has enumerable solutions $(\rho_1,\psi_1),(\rho_2,\psi_2),\ldots$, sorted as usual in ascending order of eigenvalue~\citep{garciatrillos18}.
 These eigenvalues and eigenfunctions can be used to give a spectral definition of Sobolev spaces.
@@
 
-\paragraph{Estimation.}
+\subsubsection{Estimation with spectral series regression}
 Recall the population-level spectral series estimator $\wt{f}$ defined in~\eqref{eqn:population-level_spectral_series}.
@@
 
-\paragraph{Testing.}
+\subsubsection{Testing with spectral series regression}
 In the goodness-of-fit testing problem, one asks for a test function---formally, a Borel measurable function $\phi$ that takes values in $\{0,1\}$--- which can distinguish between the hypotheses
@@
 
-\paragraph{Estimation.} PCR-LE depends on the kernel $\eta$ and two tuning parameters, the graph radius $\varepsilon$ and number of eigenvectors $K$. We will need to make some assumptions on each.
+\subsubsection{Estimation with PCR-LE} 
+PCR-LE depends on the kernel $\eta$ and two tuning parameters, the graph radius $\varepsilon$ and number of eigenvectors $K$. We will need to make some assumptions on each.
 \begin{enumerate}[label=(K\arabic*)]
@@
 Depending on the problem at hand, it may be more interesting to consider loss in $L^2(P)$ norm, for instance because this loss is intrinsically tied to the prediction error $\Ebb_{X \sim P}(\wh{f}(X) - f_0(X))^2$.
-In some preliminary analysis, we have considering applying a generic kernel smoother to the PCR-LE estimate, so as to produce a bona-fide function $T\wh{f}: \Rd \to \Reals$ for which the error $\|T\wh{f} - f_0\|_P^2$ is provably on the order of the minimax estimation rate $n^{-2/(2 + d)}$.
+In some preliminary analysis, we have considering applying a generic kernel smoother $T$ to the PCR-LE estimate, so as to produce a bona-fide function $T\wh{f}: \Rd \to \Reals$; the error $\|T\wh{f} - f_0\|_P^2$ is provably on the order of the minimax estimation rate $n^{-2/(2 + d)}$.
 Indeed, our analysis is completely independent of the structure of $\wh{f}$, and could apply to any estimator $\wh{g} \in \Reals^n$ sufficiently close to $f_0$ in empirical norm.
-We also establish that the optimal bandwidth for such a kernel smoother is meaningfully smaller than used for noisy data---which is intuitive, because we are applying it to denoised estimates---so that actually the error  $\|T\wh{f} - f_0\|_P^2$ is dominated by the in-sample mean-squared error $\|\wh{f} - f_0\|_n^2$, plus a negligible extra term.
 We intend to pursue the matter further in future work.
@@
 On the other hand, as we will see in Section~\ref{subsec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K$ is of at least the same order as the continuum eigenvalue $\rho_K$; this is essential in order to obtain a tight upper bound on the bias of $\wh{f}$.
-Finally, we set $K = \floor{(M^2n)^{d/(2 + d)}}$ (when possible) to optimally trade-off bias and variance.
-
-	% Eliminated the following chunk because I'm no longer sure what to say.
-	% In practice, one typically tunes hyper-parameters by cross-validation.
-However, because the estimator $\wh{f}$ is defined only in-sample, neither cross-validation nor any other sample-splitting technique can be used to tune parameters for PCR-LE.
-We return to this issue in Section~\ref{sec:out_of_sample}, when we propose an out-of-sample extension of $\wh{f}$.
+Finally, we set $K = \floor{(M^2n)^{d/(2 + d)}}$ (when possible) to optimally trade-off bias and variance, as is typical.
+
+	In practice, one typically tunes hyper-parameters by cross-validation.
+However, because the estimator $\wh{f}$ is defined only in-sample, cross-validation cannot be used to tune parameters for PCR-LE.
+As previously mentioned, we are currently considering methods for extrapolation of $\wh{f}$ via kernel smoothing, which should allow for cross-validation and other sample-splitting techniques.
 
@@
 
-\paragraph{Testing.} Consider the test $\varphi = \1\{\wh{T} \geq t_{a}\}$, where $t_{a}$ is the threshold
+\subsubsection{Testing with PCR-LE} 
+Consider the test $\varphi := \1\{\wh{T} \geq t_{a}\}$, where $t_{a}$ is the threshold
 \begin{equation*}
@@
 
-\paragraph{Estimation.}
+\subsubsection{Estimation with PCR-LE} 
 In order to show that $\wh{f}$ is an optimal estimator over $H_0^s(\mc{X};M)$, we will require that $\varepsilon$ be meaningfully larger than the lower bound in~\ref{asmp:parameters_estimation_fo}.
@@
 
-\paragraph{Testing.} The test $\varphi$ can adapt to the higher-order smoothness of $f_0$, when $\varepsilon$ and $K$ are chosen correctly.
+\subsubsection{Testing with PCR-LE}  The test $\varphi$ can adapt to the higher-order smoothness of $f_0$, when $\varepsilon$ and $K$ are chosen correctly.
 \begin{enumerate}[label=(P\arabic*)]
@@
 	\end{equation}
-	then $\Ebb_{f_0}[1 - \phi] \leq b$.
+	then $\Ebb_{f_0}[1 - \varphi] \leq b$.
 \end{theorem}
@@
 
-% (AG 8/30/21): Is the below chunk helpful? Or not intuitive enough.
-It is intuitively reasonable that data-dependent spectral series methods, such as PCR-LE, should have more difficulty achieving the minimax rates of convergence for testing, as opposed to estimation.
-Optimally balancing the bias and variance of the PCR-LE test requires choosing more and higher-frequency eigenvectors than in estimation.
-Recall from~\eqref{eqn:spectral_gap} that higher-frequency graph Laplacian eigenvectors---which correspond to larger eigenvalues---are meaningfully worse approximations of their population-level limits.
-Put more simply, minimax optimal testing when $d > 4$ appears to require stronger approximation properties than graph Laplacian eigenvectors can satisfy.
+At a high level, it is intuitively reasonable that PCR-LE should have more difficulty achieving the minimax rates of convergence for testing, as opposed to estimation.
+To obtain the faster rates of convergence for testing, PCR-LE must use many more eigenvectors than are necessary for estimation, including some which correspond to very large eigenvalues.
+It is known that the approximation properties of eigenvectors corresponding to large eigenvalues are very poor~\citep{burago2014,trillos2019}, and when $d > 4$ this prevents us from establishing that PCR-LE is an optimal test.
 That being said, although we suspect $\varphi$ is truly suboptimal when $d > 4$, our analysis relies on an upper bound on testing bias.
@@
 
-\subsection{Analysis}
+\subsection{Analysis of PCR-LE}
 \label{subsec:analysis}
@@
 We now outline the high-level strategy we follow when proving each of Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}.
-We analyze the estimation error of $\wh{f}$, and the testing error of $\varphi$, by first conditioning on the design points $\{X_1,\ldots,X_n\}$ and deriving \emph{design-dependent} bias and variance terms, as outlined in~\eqref{eqn:bias_variance_estimation} and~\eqref{eqn:bias_estimation}.
+We analyze the estimation error of $\wh{f}$, and the testing error of $\varphi$, by first conditioning on the design points $\{X_1,\ldots,X_n\}$ and deriving \emph{design-dependent} bias and variance terms.
 For estimation, we show that with probability at least $1 - \exp(-K)$,
@@
 \end{equation}
-These design-dependent bias-variance decompositions are reminiscent of the more classical bias-variance decompositions typical in the analysis of population-level spectral series methods (for instance\eqref{pf:spectral_series_estimation_3} and~\eqref{pf:spectral_series_test}), but different in certain key respects.
+These design-dependent bias-variance decompositions are reminiscent of the more classical bias-variance decompositions typical in the analysis of population-level spectral series methods (for instance~\eqref{pf:spectral_series_estimation_3} and~\eqref{pf:spectral_series_test}), but different in certain key respects.
 Comparing~\eqref{eqn:estimation_biasvariance} and~\eqref{eqn:testing_biasvariance} to~\eqref{pf:spectral_series_estimation_3} and~\eqref{pf:spectral_series_test}, we see that two continuum objects in the latter pair of bounds, the Sobolev norm $\|f_0\|_{\mc{H}^s(\mc{X})}^2$ and the eigenvalue $\rho_k^s$, have been replaced by graph-based analogues: the graph Sobolev seminorm $\dotp{L_{n,\varepsilon}^sf_0}{f_0}_n$ and the graph Laplacian eigenvalue $\lambda_k^s$.
@@
 
-\paragraph{Estimation.}
+\subsubsection{Estimation with PCR-LE} 
 To ensure that $\wh{f}$ is an in-sample minimax rate-optimal estimator, we choose the kernel function $\eta$, graph radius $\varepsilon$ and number of eigenvectors $K$ as in~\ref{asmp:parameters_estimation_ho}, except with ambient dimension $d$ replaced by the intrinsic dimension $m$.
@@
 
-\paragraph{Testing.}
+\subsubsection{Testing with PCR-LE}
 Likewise, to construct a minimax optimal test using $\wh{T}$, we choose $\varepsilon$ and $K$ as in~\ref{asmp:parameters_testing_fo}, except with the ambient dimension $d$ replaced by the intrinsic dimension $m$.
@@
 
-\paragraph{Analysis.}
+\subsection{Analysis}
 The high-level strategy used to prove Theorems~\ref{thm:laplacian_eigenmaps_estimation_manifold} and~\ref{thm:laplacian_eigenmaps_testing_manifold} is the same as in the flat-Euclidean setting.
@@
 	\item \emph{Manifold adaptivity}.
-An oft-recommended alternative to the population-level spectral series method considered in Section~\ref{subsec:spectral_projection} is to run OLS using eigenfunctions of a density-independent differential operator, such as the unweighted Laplacian $\Delta = \sum_{i = 1}^{d} \partial^2f/\partial x_i^2$.
-Under the conditions of Model~\ref{def:model_flat_euclidean}, such a method will indeed be rate-optimal, though the upper bounds may come with undesirably large constants if $p$ is very non-uniform.
-However, in contrast with PCR-LE, under Model~\ref{def:model_manifold} this method cannot achieve the faster minimax rates of convergence.
+When the design distribution is non-uniform, an oft-recommended alternative to population-level spectral series regression is to run OLS using eigenfunctions of a density-independent differential operator.
+As a concrete example, let $\Delta$ be the unweighted Laplacian operator on $\Rd$, $\Delta = \sum_{i = 1}^{d} \partial^2f/\partial x_i^2$.
+Denoting the eigenfunctions of $\Delta$ (under Neumann boundary conditions) by $\phi_1,\phi_2,\ldots$, and letting $\Phi \in \mathbb{R}^{n \times K}$ be the matrix with entries $\Phi_{ik} = \phi_k(X_i)$ and columns $\Phi_1,\ldots,\Phi_K$,  one could compute an estimator by solving the following OLS problem:
+	\begin{equation*}
+	\minimize_{f \in \mathrm{span}\{\Phi_{1},\ldots,\Phi_{K}\}} \|{\bf Y} - f\|_n^2.
+	\end{equation*}
+	Unlike with spectral series regression, this approach can produce reasonable estimates even when the sampled eigenfunctions $(\phi_k(X_1),\ldots,\phi_k(X_n)) \in \Reals^n$ are not approximately orthogonal.
+Indeed, under the conditions of Model~\ref{def:model_flat_euclidean}, such a method will in fact be minimax rate-optimal, though the upper bounds may come with undesirably large constants if $p$ is very non-uniform.
+However under Model~\ref{def:model_manifold}, this method cannot achieve the faster minimax rates of convergence---which depend only on the intrinsic dimension $m$---and may even be inconsistent.
+This is because the eigenfunctions $\phi_k$ have no intrinsic relationship to the Sobolev space $H^s(\mc{X})$ except when $\mc{X}$ is a full-dimension set in $\Rd$.
+In contrast, PCR-LE uses features which are empirical approximations to eigenfunctions $\psi_k$ of the density-weighted Laplace-Beltrami operator $\Delta_P$.
+The eigenfunctions of $\Delta_P$ are thus adapted to the manifold $\mc{X}$, and PCR-LE is consistent and in certain cases optimal, as we have shown.
 	\item \emph{Density adaptivity}.

### 2021-12-22 code/estimators.R
Minor changes to code base.

@@ -80,9 +80,16 @@ make_laplacian_smoothing <- function(theta)
   r <- theta[["r"]]
   rho <- theta[["rho"]]
   labeled <- theta[["labeled"]]
+  kernel <- theta[["kernel"]]
   laplacian_smoothing <- function(Y,X){
+    
     # Build G_{n,r} over X.
-    G <- neighborhood_graph(X,r)
+    if(!is.null(kernel))
+    {
+      G <- neighborhood_graph(X,r,kernel)
+    } else{
+      G <- neighborhood_graph(X,r)
+    }
     
     # Get L.
     L <- Laplacian(G)

### 2021-12-22 code/graph.R

@@ -41,6 +41,7 @@ knn_to_neighborhood_graph <- function(X, r, k_max, X_new = X,loop, kernel)
     
     r_neighbors_list <- as.matrix(melt(rneighbors)[,-2])
     r_neighbors_list <- r_neighbors_list[rowSums(is.na(r_neighbors_list)) == 0,]
+    r_neighbors_list <- r_neighbors_list[,c(2,1)]
     
     dist_to_neighbors_vec <- c(dist_to_neighbors)
     dist_to_neighbors_vec <- dist_to_neighbors_vec[!is.na(dist_to_neighbors_vec)]

### 2021-12-22 code/hat_matrices.R

@@ -2,11 +2,16 @@ make_laplacian_smoothing_hat <- function(theta)
 {
   r <- theta[["r"]]
   rho <- theta[["rho"]]
-  kernel <- theta[["kernel"]]
+  kernel <- theta[["kernel"]][[1]]
   laplacian_smoothing <- function(X){
     
     # Build G_{n,r} over X.
-    G <- neighborhood_graph(X,r)
+    if(!is.null(kernel))
+    {
+      G <- neighborhood_graph(X,r,kernel = kernel)
+    } else{
+      G <- neighborhood_graph(X,r)
+    }
     
     # Get L.
     L <- Laplacian(G)

### 2021-12-22 code/misc/example_for_muri_presentation.R

@@ -1,0 +1,120 @@
+#------------------------------#
+# Script to create warmup example
+#------------------------------#
+setwd("~/Statistics/graph_regression/code")
+source("graph.R")
+source("sample.R")
+library(Matrix)
+library(magrittr)
+library(RANN)
+library(reshape2)
+library(RSpectra)
+setwd("~/Statistics/graph_regression/talks/laplacian_smoothing/MURI/figures/")
+
+# Data
+set.seed(4444)
+n <- 200
+d <- 2
+sampler <- function(n)
+{
+  sample_x <- make_sample_gaussian_mixture(d = 1, mu = c(-1,1), pi = c(.5,.5),sigma = c(.5,.5),truncate = T)
+  x <- sample_x(n)
+  y <- sapply(x, FUN = function(x_i){runif(1,- (sqrt(x_i^4*(1 - x_i^2)) + .2), sqrt(x_i^4*(1 - x_i^2)) + .2)}) # dumbbell
+  X <- matrix(c(x,y),ncol = 2)
+}
+X <- sampler(n)
+f0 <- function(x){2*x[1]}
+Y <- apply(X,1,f0) + rnorm(n,0,.5)
+
+# Compute graph, keep largest connected component.
+radius <- c(.15)
+kernel_fxn <- function(s){exp(-2*s^2)}
+G <- neighborhood_graph(X,radius,kernel = kernel_fxn)
+X <- X[which(rowSums(G) > 0),]
+Y <- Y[which(rowSums(G) > 0)]
+G <- neighborhood_graph(X,radius,kernel = kernel_fxn) # one connected component
+adj_list <- which(G > 0, arr.ind = T)
+adj_list <- cbind(adj_list, G[adj_list])
+
+# 1. Plot X and Y
+coul <- colorRampPalette(c("blue","red"))
+coul <- coul(25)
+colors <- coul[cut(Y,25)]
+pdf("warmup_example_1.pdf")
+par(pty="s",mar = c(1,1,1,1))
+plot(X[,1],X[,2], 
+     ylim = c(-1.1,1.1),xlim = c(-1.1,1.1),
+     axes = F,xlab = "",ylab = "",type = "n")
+points(X[,1],X[,2],pch = 21, bg = colors, col = "black",cex = 2)
+dev.off()
+
+# 2. Plot neighborhood graph.
+
+# Add more colors to this palette :
[…]

### 2021-12-22 code/plot_methods.R

@@ -325,6 +325,32 @@ plot_eigenvectors <- function(V,X,Ks = 1:ncol(V),ncolors = 25,cex.pt = 3){
   }
 }
 
+plot_equivalent_kernel <- function(H,X, plot_X, theta,
+                                  labeled = T,
+                                  colors = rep("black",length(plot_X)),
+                                  rug = T,
+                                  title = NULL,
+                                  xlim = NULL)
+{
+  if(d == 1)
+  {
+    plot_equivalent_kernel_1d(H,X, plot_X, theta,
+                              labeled = labeled,
+                              colors = colors,
+                              title = title,
+                              rug = rug,
+                              xlim = xlim)
+  } else if(d == 2)
+  {
+    plot_equivalent_kernel_2d(H,X, plot_X, theta,
+                              labeled = labeled,
+                              colors = colors,
+                              title = title,
+                              wireframe = T,
+                              xlim = xlim)
+  }
+}
+
 # Plot equivalent kernels.
 plot_equivalent_kernel_1d <- function(H,X, plot_X, theta,
                                       labeled = T,
@@ -334,14 +360,14 @@ plot_equivalent_kernel_1d <- function(H,X, plot_X, theta,
                                       xlim = NULL)
 {
   plot_indx <- apply(plot_X,1,FUN = function(x){apply(X,1,FUN = function(z){sum(z - x)^2}) %>% which.min()})
-  plot_mat <- H[plot_indx,]
+  plot_mat <- H[plot_indx,,drop = F]
   
   if(is.null(title))
   {
-    if(all(names(theta) == c("r","rho","kernel"))){
+    if(all(c("r","rho") %in% names(theta))){
       title <- paste0("Radius: ", round(theta$r,2), 
                       ". Rho: ",round(theta$rho,2),".")
-    } else if(all(names(theta) == c("r","K","kernel"))){
+    } else if(all(c("r","K") %in% names(theta))){
       title <- paste0("Radius: ", round(theta$r,2), 
                       ". K: ",K,".")
     }
@@ -377,6 +403,52 @@ plot_equivalent_kernel_1d <- function(H,X, plot_X, theta,
   }
   if(rug){rug(X, lwd = .01)}
   grid(lwd = 2)
+}
+
[…]

### 2021-12-22 manuscripts/laplacian_eigenmaps/arxiv/laplacian_eigenmaps_regression.tex
Added cover letter for Inf and Inf.

@@
 Under sufficient smoothness conditions on the design density $p$, PCR-LE achieves the optimal rates for both estimation (where the optimal rate in squared $L^2$ norm is known to be $n^{-2s/(2s + d)}$) and goodness-of-fit testing ($n^{-4s/(4s + d)}$).
-We also show that PCR-LE is \emph{manifold adaptive}: that is, we consider the situation where $\mc{X}$ is a manifold of small intrinsic dimension $m$, and give upper bounds establishing that PCR-LE achieves the faster minimax estimation ($n^{-2s/(2s + m)}$) and testing ($n^{-4s/(4s + m)}$) rates of convergence.
+We also show that PCR-LE is \emph{manifold adaptive}: that is, we consider the situation where the design is supported on a manifold of small intrinsic dimension $m$, and give upper bounds establishing that PCR-LE achieves the faster minimax estimation ($n^{-2s/(2s + m)}$) and testing ($n^{-4s/(4s + m)}$) rates of convergence.
 Interestingly, these rates are almost always much faster than the known rates of convergence of graph Laplacian eigenvectors to their population-level limits; in other words, for this problem regression with estimated features appears to be much easier, statistically speaking, than estimating the features itself.

### 2021-12-22 manuscripts/laplacian_eigenmaps/information_and_inference/cover_letter.tex
Added cover letter for Inf and Inf.

\documentclass{article}
\usepackage{enumerate}

\begin{document}

Dear Editors,

\vspace{.2 in}

Please find our attached submission titled: Minimax Optimal Regression over Sobolev Spaces via Laplacian Eigenmaps on Neighborhood Graphs. We analyze Principal Components Regression with Laplacian Eigenmaps, a spectral projection approach to nonparametric regression that projects observed responses onto the span of eigenvectors of a neighborhood graph Laplacian. We analyze this method in a classical random design regression model; in this model, we establish upper bounds showing that the method can optimally take advantage of Sobolev smoothness in the underlying regression function, and low intrinsic dimension in the design data manifold. Interestingly, these upper bounds imply that our estimator converges at rates far faster than the rates at which graph Laplacian eigenvectors converge to their continuum limits.

For completeness, we also derive some minimax rates for Sobolev classes, in the less well-studied subcritical regime where such classes contain discontinuous functions. However, we stress that our main contribution is the analysis of Principal Components Regression with Laplacian Eigenmaps, rather than novel minimax theory. Thank you for your consideration of this manuscript.
\vspace{.2 in}

Sincerely,

\vspace{.1 in}
Alden Green, Sivaraman Balakrishnan, and Ryan Tibshirani

\end{document}

### 2021-12-22 manuscripts/laplacian_eigenmaps/main_text.tex
Added cover letter for Inf and Inf.

@@
 \end{enumerate} 
-A natural way to use LE is by taking the vectors $\{(v_{1,i},\ldots,v_{K,i})\}_{i = 1}^{n}$ to be features in a downstream regression algorithm.
+A natural way to use LE is by taking the collection of vectors $\{(v_{1,i},\ldots,v_{K,i})\}_{i = 1}^{n}$ to be features in a downstream regression algorithm.
 In this paper, we study a simple method along these lines: Principal Components Regression with Laplacian-Eigenmaps (PCR-LE), a method for nonparametric regression which operates by running ordinary least squares (OLS) using the features output by LE.
@@
 Spectral series regression is intrinsically linked with Sobolev spaces.
-That is because under appropriate boundary conditions, a unit ball in the order-$s$ Sobolev space consists of functions $f = \sum_{k} a_k \psi_k \in L^2(\mc{X})$ for which the generalized Fourier coefficients $\{a_k\}_{k = 1}^{\infty}$ satisfy the decay condition $\sum_{k} a_k^2 \rho_k^s \leq C$ (See Section~\ref{subsec:spectral_projection} for more details).
+That is because under appropriate boundary conditions, a ball in the order-$s$ Sobolev space consists of functions $f = \sum_{k} a_k \psi_k \in L^2(\mc{X})$ for which the generalized Fourier coefficients $\{a_k\}_{k = 1}^{\infty}$ satisfy the decay condition $\sum_{k} a_k^2 \rho_k^s \leq C$ (See Section~\ref{subsec:spectral_projection} for more details).
 This decay condition justifies the truncated series estimator \eqref{eqn:population-level_spectral_series}, since it means the truncation will incur only a limited amount of bias for any $f_0 \in H^s(\mc{X})$.
@@
 Viewed in this light, a major advantage of PCR-LE is that it operates without needing knowledge of the design distribution $P$.
-This is an advantage because in our context $P$ is treated as an unknown and potentially complex distribution, and for example can be highly non-uniform, have a complicated support which may be a submanifold of $\Rd$, or both.
+This is an advantage because in our context $P$ is an unknown and potentially complex distribution: for example, it can be highly non-uniform, have a complicated support which may be a submanifold of $\Rd$, or both.
 In contrast, spectral series regression relies on diagonalizing the density-weighted Laplacian $\Delta_P$, and in our context must be viewed as an oracle method; to emphasize this we henceforth refer to the estimator defined in~\eqref{eqn:population-level_spectral_series} as \emph{population-level spectral series regression}.
@@
 
-\paragraph{Perspective: Regression with Empirical Features vs. Feature Estimation.}
+\paragraph{Perspective: regression error versus feature reconstruction.}
 We now pause for a moment, to emphasize that in a certain respect the aforementioned rates of convergence for PCR-LE are quite surprising.
@@
 Crucially, our analysis \emph{does not} work in this way.
-This is important because all known upper bounds on the rates at which $v_k \to \psi_K$ as $n \to \infty$ are much slower than the minimax rates for regression over Sobolev classes.
-For instance, the best currently known upper bound on the empirical $L^2$ error $\frac{1}{n}\sum_{i = 1}^{n}(\sqrt{n} v_{k,i} - \psi_k(X_i))^2$ is only on the order of $n^{-2/(4 + d)}$~\cite{cheng2021}, which is slower than the minimax estimation rate over $H^s(\mc{X})$ for any $s \in \mathbb{N}, s \geq 1$.\footnote{To make matters worse, PCR-LE, when deployed optimally, does not use a single eigenvector $v_k$ for a fixed index $k \in \mathbb{N}$, but rather many eigenvectors $v_1,\ldots,v_K$ with $K$ growing in $n$.
+This is important because all known upper bounds on the rates at which $v_k \to \psi_k$ as $n \to \infty$ are much slower than the minimax rates for regression over Sobolev classes.
+For instance, the best currently known upper bound on the empirical $L^2$ error $\frac{1}{n}\sum_{i = 1}^{n}(\sqrt{n} v_{k,i} - \psi_k(X_i))^2$ is only on the order of $n^{-2/(4 + d)}$~\citep{cheng2021}, which is slower than the minimax estimation rate over $H^s(\mc{X})$ for any $s \in \mathbb{N}, s \geq 1$.\footnote{To make matters worse, PCR-LE, when deployed optimally, does not use a single eigenvector $v_k$ for a fixed index $k \in \mathbb{N}$, but rather many eigenvectors $v_1,\ldots,v_K$ with $K$ growing in $n$.
 As $K$ grows larger, the rate at which $v_K \to \psi_K$ gets slower, since the population-level object being estimated is less regular; see~\citep{burago2014,trillos2019}.} Although this upper bound may not reflect the true rate of convergence of graph Laplacian eigenvectors---this is still an active area of research, and no lower bounds are known---it seems very unlikely that the true rate matches the minimax estimation rate $n^{-2s/(2s + d)}$, which after all approaches the dimension-free rate $1/n$ for large values of $s$.
@@
 Ultimately our upper bound on the error of PCR-LE is determined entirely by a pair of graph functionals: the quadratic form $f_0^{\top}L^s f_0$, and the graph Laplacian eigenvalue $\lambda_{K + 1}$.
-This brings a couple of advantages.
-First, it eliminates the need to analyze convergence of eigenvectors to eigenfunctions, which is critical in order to get sufficiently fast rates of convergence for PCR-LE, as we have already explained.
+This brings a couple of advantages:
+\begin{itemize}
+	\item First, it eliminates the need to analyze convergence of eigenvectors to eigenfunctions, which is critical in order to get sufficiently fast rates of convergence for PCR-LE, as we have already explained.
 Instead, we only have to consider these two graph functionals, both of which are known to converge at faster rates than graph Laplacian eigenvectors.
-Second, in order to obtain upper bounds on $\|\wh{f} - f_0\|_n^2$ we do not require that these graph functionals themselves converge to population-level limits, but only that they be stochastically bounded on the right order, which is a much weaker requirement.
+	\item Second, in order to obtain upper bounds on $\|\wh{f} - f_0\|_n^2$ we do not require that these graph functionals themselves converge to population-level limits, but only that they be stochastically bounded on the right order.
+The latter is a much weaker requirement.
+\end{itemize}
 To derive our upper bounds on the error of PCR-LE, we directly analyze the quadratic form $f_0^{\top}L^s f_0$ and the eigenvalue $\lambda_{K + 1}$, using some existing results as well as deriving some new ones which may be of independent interest.
@@
 In this supervised setting, there has been relatively little work regarding \emph{random design} regression with neighborhood graph Laplacians .
-Aside from our own work on Laplacian smoothing, summarized above, we highlight two other related papers: \citet{lee2016}, who analyze a variant of PCR-LE, but derive suboptimal rates of convergence, and \citet{trillos2020}, who study Laplacian smoothing and establish the uniform upper bound $\max_{i = 1,\ldots,n}|\wc{f}(X_i) - f_0(X_i)| \leq C n^{-2/(2 + d)}$ under the assumption $f_0 \in C^2(\mc{X})$, which is slower than the minimax rate $n^{-4/(4 + d)}$ for this class.
+Aside from our own work on Laplacian smoothing, summarized above, we highlight two other related papers: \citet{lee2016}, who analyze a variant of PCR-LE, but derive suboptimal rates of convergence, and \citet{trillos2020}, who study Laplacian smoothing and establish the uniform upper bound $\max_{i = 1,\ldots,n}|\wc{f}(X_i) - f_0(X_i)| \leq C n^{-2/(2 + d)}$ under the assumption $f_0 \in C^2(\mc{X})$, which is slower than the minimax rate $n^{-4/(4 + d)}$ for this function class.
 
@@
 However, they are unusual in that we treat both the case where $s < d/2$ and thus the Sobolev space $H^s(\mc{X})$ does not continuously embed into $C(\mc{X})$, and the case where $P$ is not the uniform distribution over the unit cube.
-The upper bounds given in this section serve two purposes: first, to clarify what the rates are in these less-typically studied settings; second, to show that population-level spectral series regression can obtain the optimal rate.
+The upper bounds given in this section serve two purposes: first, to clarify what the rates are in these less-typically studied settings; second, to show that even in this general setting, population-level spectral series regression can always obtain the optimal rates.
 This latter point is important since the method we focus on for the most part, PCR-LE, is an empirical approximation to population-level spectral series regression.
@@
 	\end{equation}
-	then the Type II error is upper bounded, $\Ebb_{f_0}[1 - \phi] \leq b$.
+	then the Type II error is upper bounded, $\Ebb_{f_0}[1 - \wt{\varphi}] \leq b$.
 \end{proposition}
-Assuming again that $n^{-1/2} \lesssim M$, the upper bound in~\eqref{eqn:spectral_series_testing} is $M^2(M^2n)^{-4s/(4s + d)}$, matching the usual minimax critical radius over Sobolev spaces (see e.g.~\cite{guerre02,ingster2009,ingster2012}.
-Specifically, \cite{ingster2009} show that the minimax squared critical radius is on the order of $n^{-4s/(4s + d)}$ when $M = 1$, and simple alterations of their analysis imply the rate $M^2(M^2n)^{-4s/(4s + d)}$ for general $M$.)
-
-On the other hand, when $s \leq d/4$ the minimax regression testing rates over $H^s(\mc{X})$ are not known.
+Assuming again that $n^{-1/2} \lesssim M$, the right hand side of~\eqref{eqn:spectral_series_testing} is $M^2(M^2n)^{-4s/(4s + d)}$, matching the usual minimax critical radius over Sobolev spaces (see e.g.~\cite{guerre02,ingster2009,ingster2012}.
+Specifically, \cite{ingster2009} show that the minimax squared critical radius is on the order of $n^{-4s/(4s + d)}$ when $M = 1$, and simple alterations of their analysis imply the rate $M^2(M^2n)^{-4s/(4s + d)}$ for general $M$.) On the other hand, when $s \leq d/4$ the minimax regression testing rates over $H^s(\mc{X})$ are not known.
 If one explicitly assumes $f_0 \in L^4(\mc{X};1)$---note that $H^s(\mc{X})$ does not continuously embed into $L^4(\mc{X})$ when $s \leq d/4$---then the minimax critical radius for regression testing is on the order of $n^{-1/2}$ \citep{guerre02}, and is achieved by a test using the naive statistic $\|{\bf Y}\|_n^2$.
@@
 Depending on the problem at hand, it may be more interesting to consider loss in $L^2(P)$ norm, for instance because this loss is intrinsically tied to the prediction error $\Ebb_{X \sim P}(\wh{f}(X) - f_0(X))^2$.
-In some preliminary analysis, we have considering applying a generic kernel smoother $T$ to the PCR-LE estimate, so as to produce a bona-fide function $T\wh{f}: \Rd \to \Reals$; the error $\|T\wh{f} - f_0\|_P^2$ is provably on the order of the minimax estimation rate $n^{-2/(2 + d)}$.
+In some preliminary analysis, we have considered applying a generic kernel smoother $T$ to the PCR-LE estimate, so as to produce a bona-fide function $T\wh{f}: \Rd \to \Reals$; the error $\|T\wh{f} - f_0\|_P^2$ is provably on the order of the minimax estimation rate $n^{-2/(2 + d)}$.
 Indeed, our analysis is completely independent of the structure of $\wh{f}$, and could apply to any estimator $\wh{g} \in \Reals^n$ sufficiently close to $f_0$ in empirical norm.
@@
 At a high level, it is intuitively reasonable that PCR-LE should have more difficulty achieving the minimax rates of convergence for testing, as opposed to estimation.
-To obtain the faster rates of convergence for testing, PCR-LE must use many more eigenvectors than are necessary for estimation, including some which correspond to very large eigenvalues.
+To obtain the faster rates of convergence for testing, PCR-LE must use many more eigenvectors than are necessary for estimation, including some eigenvectors which correspond to very large eigenvalues.
 It is known that the approximation properties of eigenvectors corresponding to large eigenvalues are very poor~\citep{burago2014,trillos2019}, and when $d > 4$ this prevents us from establishing that PCR-LE is an optimal test.
@@
 However under Model~\ref{def:model_manifold}, this method cannot achieve the faster minimax rates of convergence---which depend only on the intrinsic dimension $m$---and may even be inconsistent.
-This is because the eigenfunctions $\phi_k$ have no intrinsic relationship to the Sobolev space $H^s(\mc{X})$ except when $\mc{X}$ is a full-dimension set in $\Rd$.
+This is because the eigenfunctions $\phi_k$ have no underlying relationship to the Sobolev space $H^s(\mc{X})$ except when $\mc{X}$ is a full-dimension set in $\Rd$.
 In contrast, PCR-LE uses features which are empirical approximations to eigenfunctions $\psi_k$ of the density-weighted Laplace-Beltrami operator $\Delta_P$.
-The eigenfunctions of $\Delta_P$ are thus adapted to the manifold $\mc{X}$, and PCR-LE is consistent and in certain cases optimal, as we have shown.
+The eigenfunctions of $\Delta_P$ are appropriately adapted to the geometry of the manifold $\mc{X}$, and as a result PCR-LE is consistent and in certain cases minimax optimal, as we have shown.
 	\item \emph{Density adaptivity}.
-In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we give a simple example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that the expected in-sample mean squared error of PCR-LE is smaller than that of either kernel smoothing or least squares using eigenfunctions of $\Delta$.
+In Appendix~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we give a simple univariate example of a sequence of densities and regression functions $\{(p^{(n)}, f_0^{(n)}: n \in \mathbb{N}\}$ such that the expected in-sample mean squared error of PCR-LE is smaller than that of either kernel smoothing or least squares using eigenfunctions of $\Delta$.
 This is possible because PCR-LE induces a completely different bias than these latter two methods.

### 2021-12-22 notes/misc/local_weyls_law.tex
Some miscellaneous notes.

--- 2021-03-08 notes/weekly_notes/12.24.20.NotesForWeek.tex
@@
 \begin{document}
-\title{Note on: Out-of-Sample Error for Laplacian Eigenmaps}
+\title{Note on: Local Weyl's Law}
 \author{Alden Green}
@@
 
-The graph Laplacian eigenmaps estimator $\wh{\theta}_{\LE}$ is a least-squares estimator, where a response vector $Y = (Y_1,\ldots,Y_n)$ is projected onto the span of the first $K$ eigenvectors of a graph Laplacian matrix.
-It is defined only in-sample, whereas out-of-sample error is more interesting in many applications.
-There is no commonly agreed upon way in the literature to extend $\wh{\theta}_{\LE}$, or indeed any type of Laplacian estimator.
-In this note, I discuss three methods for doing so: least-squares regression of $\wh{\theta}_{\LE}$ onto an a priori defined set of features, Nadaraya-Watson kernel smoothing, treating $\wh{\theta}_{\LE}$ as the response, and Nystr\"{o}m extension of Laplacian eigenvectors.
-These methods are roughly ordered from least to most preferable, but also (as these things go) easiest to hardest to analyze.
-I begin in Section~\ref{sec:least_squares_arbitrary_design_and_features} by reviewing known bounds on the out-of-sample error of a (generic) least-squares estimator; these bounds will be useful for the analysis of both the first and third methods.
+Let $\Delta = -\sum_{i = 1}^{d} \frac{\partial^2}{\partial x_i^2}$ be the continuum Laplacian operator in $\Rd$.
+Let $\mc{X}$ be a connected open bounded set in $\Rd$, with $C^{1,1}$ boundary.
+The eigenvector equation with Neumann boundary conditions is
+\begin{equation}
+\label{eqn:eigenvector}
+\Delta \psi = \lambda \psi,\quad \frac{\partial}{\partial {\bf n}}\psi = 0~~\textrm{on $\partial \mc{X}$.}
+\end{equation}
+The operator $\Delta$ has a discrete spectrum, by which we mean that there exist enumerable pairs $(\lambda,\psi)$ which satisfy~\eqref{eqn:eigenvector}.
+We enumerate them $(\lambda_1,\psi_1),(\lambda_2,\psi_2),\ldots$ in increasing order of eigenvalue, $\lambda_1 \leq \lambda_2 \leq \ldots$.
 
-\section{Out-of-sample error for least-squares estimators: arbitrary design and features}
-\label{sec:least_squares_arbitrary_design_and_features}
-This section largely summarizes results from a paper by Daniel Hsu on random design regression. We start with some general setup that will apply to everything that follows.
+Note that $(\psi_k)_{k}$ form an orthonormal basis of $L^2(\mc{X})$.
+Thus the spectral function $E_{\lambda}u := \sum_{k: \lambda_k \leq \lambda} \dotp{u}{\psi_k} \psi_k$ is the projection of a given $u \in L^2(\mc{X})$ onto the ``bottom'' part of the spectrum.
+The following estimate on the spectral function is due to H\"{o}rmander, Theorem 17.5.3:
+\begin{equation}
+\label{eqn:local_weyl_law}
+\sup_{x \in \mc{X}} |E_{\lambda}(x)|^2 \leq C \lambda^{d/2} \quad \textrm{for all $K \in \mathbb{N}$.}
+\end{equation}
+I prove~\eqref{eqn:local_weyl_law}, following the proof of H\"{o}rmander but giving some extra exposition where I didn't initially understand his reasoning.
+First, I make some useful notes about the result.
+\begin{itemize}
+	\item The statement~\eqref{eqn:local_weyl_law} also holds when $\Delta$ is replaced by a symmetric second order positive differential operator,
+	\begin{equation*}
+	\mc{L} :=  -\sum_{i,j = 1}^{d} g^{ij}\frac{\partial^2}{\partial x_ix_j} + \sum_{j = 1}^{d} b^j \frac{\partial}{\partial x_j} + c
+	\end{equation*}
+	with coefficients $g$ and $b$ in $C^{\infty}(\mc{X})$. This includes the case of the weighted Laplace-Beltrami operator $\Delta_P$ when $P$ admits a smooth density $p$ over $\mc{X}$.
+	\item Let $N(\lambda) = \#\{k: \lambda_k \leq \lambda\}$. Using the trace trick, we have that
+	\begin{equation*}
+	N(\lambda) = \sharp\int_{\mc{X}} |E_\lambda(x)|^2 \,dx \leq C\lambda^{d/2}
+	\end{equation*}
+	implying that $\lambda_K \gtrsim K^{2/d}$. This is one-half of the Weyl's law asymptotic scaling $\lambda_K \asymp K^{2/d}$; in fact it is only half needed to establish 
+	\item 
+\end{itemize}
 
-Suppose we observe $(X_1,\wh{\theta}_1),\ldots,(X_n,\wh{\theta}_n)$, where $X_1,\ldots,X_n \in \Reals^d$ and $\wh{\theta}_{1:n} = (\wh{\theta}_1,\ldots,\wh{\theta}_n) \in \Reals^n$.
-Our goal is to use this observed data to construct an estimator $\wh{f}$ which is on average close to $f^{\ast}: \Omega \to \Reals$ on the domain $\Omega$ (for instance, $\Omega = [0,1]^d$.) Specifically, we will consider the following least-squares estimator,
-\begin{equation*}
-\wh{\beta} = \argmin \| \wh{\theta}_{1:n} - \Phi_n^{\top} \beta \|_n^2 = (\Phi_n^\top \Phi_n)^{-1}\Phi_n^{\top} \wh{\theta}_{1:n},~~ \wh{f} = \sum_{k = 1}^{\kappa} \wh{\beta}_k \varphi_k
-\end{equation*}
-where $\varphi_1,\ldots,\varphi_{K}$ are each elements of $\Leb^2(\Omega)$, and $\Phi_n \in \Reals^{n \times K}$ has entries $(\Phi_{n})_{ik} = \varphi_k(X_i)$.
-If $\wh{\theta}_{1:n}$ lies in the span of the columns of $\Phi_n$, then $\wh{f}$ is a interpolant of $\wh{\theta}_{1:n}$; this applies to our third method, Nystr\"{o}m extension of Laplacian eigenvectors, but not to our first method.
-We will measure error in the $\Leb^2(P)$ norm $\norm{\wh{f} - f^{\ast}}_P^2 = \int_{\Omega} (\wh{f}(x) - f^{\ast}(x))^2 \,dP(x)$, where $P$ is a probability measure supported on domain $\Omega$.
-For this section, we assume no relationship between $P$ and $X_1,\ldots,X_n$.
+and show that it implies the asymptotic estimate $\lambda_K \gtrsim K^{2/d}$, which is one-half of the Weyl's Law estimate.
+Note that the other half of Weyl's Law, $\lambda_K \lesssim K^{2/d}$, along with~\eqref{eqn:local_weyl_law}, implies that
+\begin{equation}
+\label{eqn:local_weyl_law2}
+\sup_{x \in \mc{X}} |E_{\lambda_K}(x)|^2 \leq C K \quad \textrm{for all $K \in \mathbb{N}$.}
+\end{equation}
+~\eqref{eqn:local_weyl_law2} is how we state the result in our Laplacian eigenmaps paper.
 
-We begin by decomposing this norm into several modular terms, which capture the different types of error intrinsic to the problem. To this end, let
-\begin{align*}
-\wt{\beta} & = \argmin_{\beta \in \Reals^K} \|f^{\ast}_{1:n} - \Phi_n^{\top} \beta \|_n^2 = (\Phi_n^\top \Phi_n)^{-1}\Phi_n^{\top} {f}^{\ast}_{1:n},&& \wt{f}_K = \sum_{k = 1}^{K} \wt{\beta}_k \varphi_k \\
-\beta^{\star} & = \argmin_{\beta \in \Reals^K} \|f^{\ast} - \sum_{k = 1}^{K} \beta_k \varphi_k \|_P^2,&& f_K^{\ast} = \sum_{k = 1}^{K} \beta_k^{\ast} \varphi_k
-\end{align*}
-and let $f_{\perp}^{\ast} = f^{\ast} - f_{K}^{\ast}$. We decompose $\norm{\wh{f} - f^{\ast}}_P^2$ as follows:
-\begin{equation*}
-\norm{\wh{f} - f^{\ast}}_P^2 \leq 3\Bigl(\underbrace{\norm{\wh{f} - \wt{f}_K}_P^2}_{\textrm{estimation error}} + \underbrace{\norm{\wt{f}_K - f_K^{\ast}}_P^2}_{\textrm{sampling error}} + \underbrace{\norm{f_K^{\ast} - f^{\ast}}_P^2}_{\textrm{truncation error}}\Bigr)
-\end{equation*}
-This leaves us with three separate kinds of error, and we now upper bound the first two.
-
-\paragraph{Notation.} Let $e_{1:n} = \wh{\theta}_{1:n} - {f}^{\ast}_{1:n}$.
-Let~$\Sigma \in \Reals^{K \times K}$ be the second moment matrix of $\varphi_1,\ldots,\varphi_K$ with $(k,\ell)$-th entry $\Sigma_{k\ell} = \Ebb_{X \sim P}\bigl[\varphi_k(X) \varphi_{\ell}(X)\bigr]$.
-Likewise let $\wh{\Sigma} = \Phi_n^{\top} \Phi_n/n$ be the empirical second moment matrix.
-For a matrix $A \in \Reals^{K \times K}$, let $\|A\|$ denote the operator norm of $A$.
-For a probability distribution $Q$ on $\Omega$, let $d_{\min}(Q) = \inf_{x \in \Omega} \int K(\|z - x\|/r) \,dQ(z)$ be the minimum degree of $Q$.
-
-\paragraph{Estimation error.}
-Noting that
-\begin{equation*}
-~~\wh{\beta} - \wt{\beta} = \frac{1}{n}\wh{\Sigma}^{-1} \Phi_n^{\top} e_{1:n},
-\end{equation*}
-the estimation error satisfies the inequality,
-\begin{align*}
-\norm{\wh{f} - \wt{f}_K}_P^2 & = (\wh{\beta} - \wt{\beta})^{\top} \Sigma (\wh{\beta} - \wt{\beta}) \\
-& = \frac{1}{n^2} e_{1:n}^{\top} \Phi_n \wh{\Sigma}^{-1} \Sigma \wh{\Sigma}^{-1} \Phi_n^{\top} e_{1:n} \\
-& \leq \|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\| \cdot \|e_{1:n}\|_n^2.
-\end{align*}
-with the inequality following since $\|AB\| = \|BA\|$.
-
-\paragraph{Sampling error.}
-Multiplying $\wt{\beta} - \beta^{\star}$ on the left by $\Sigma^{1/2}$, we have
-\begin{equation*}
-\Sigma^{1/2}(\wt{\beta} - \beta^{\ast}) =  \Sigma^{1/2} \wh{\Sigma}^{-1} \bigl( \frac{1}{n} \Phi_n^{\top} f_{1:n}^{\ast} - \frac{1}{n} \Phi_n^{\top} \Phi_n \beta^{\ast} \bigr) = \Sigma^{1/2} \wh{\Sigma}^{-1} \frac{1}{n} \Phi_n^{\top}  \bigl(\underbrace{f_{1:n}^{\ast} - \Phi_n \beta^{\ast}}_{ = (f_{\perp}^{\ast})_{1:n}} \bigr) = \frac{1}{n}\Sigma^{1/2} \wh{\Sigma}^{-1} \Phi_n^{\top} (f_{\perp}^{\ast})_{1:n}.
-\end{equation*}
-Therefore
-\begin{equation*}
-\|\wt{f}_K - f_K^{\ast}\|_P^2  = \|\wt{\beta} - \beta^{\ast}\|_{\Sigma}^2 \leq \|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\| \cdot \Bigl\|f_{\perp}^{\ast}\|_n^2
-\end{equation*}
-
-The results of this section are summarized by Lemma~\ref{lem:l2p_error_fixed_design}. 
-\begin{lemma}
-	\label{lem:l2p_error_fixed_design}
-	Suppose $\wh{\Sigma} \succeq 0$. Then the $\Leb^2(P)$ error $\|\wh{f} - f^{\ast}\|_P^2$ is upper bounded as follows:
-	\begin{equation}
-	\label{eqn:l2p_error_fixed_design}
-	\|\wh{f} - f^{\ast}\|_P^2 \leq 3\Bigl(\|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\| \cdot \|\wh{\theta}_{1:n} - {f}^{\ast}_{1:n}\|_n^2 + \|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\| \cdot \|f_{\perp}^{\ast}\|_n^2 + \|f_{\perp}^{\ast}\|_P^2\Bigr).
-	\end{equation}
-	If $\wh{\Sigma} = I$ then
-	\begin{equation}
-	\label{eqn:l2p_error_fixed_design_identity}
-	\|\wh{f} - f^{\ast}\|_P^2 \leq 3\Bigl(\|\Sigma\| \cdot \|\wh{\theta}_{1:n} - {f}^{\ast}_{1:n}\|_n^2 + \|\Sigma\| \cdot \|f_{\perp}^{\ast}\|_n^2 + \|f_{\perp}^{\ast}\|_P^2\Bigr).
-	\end{equation}
-\end{lemma}
-(Basically) no assumptions so far have been made on $X_1,\ldots,X_n$ or $\varphi_1,\ldots,\varphi_K$.
-In the standard setup $X_1,\ldots,X_n$ are randomly sampled from $P$, and $\varphi_1,\ldots,\varphi_K$ are a priori determined.
-Then matrix Bernstein's inequality gives an upper bound on $\|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\|$, standard Bernstein's inequality gives an upper bound on $\|f_{\perp}^{\ast}\|_n^2 - \|f_{\perp}^{\ast}\|_P^2$, and we obtain from~\eqref{eqn:l2p_error_fixed_design} an upper bound that depends on $\|e_{1:n}\|_n^2$ and $\|f_{\perp}^{\ast}\|_P^2$.
-
-If $\varphi_1,\ldots,\varphi_{K}$ are chosen in a data-dependent manner, then the analysis is plainly quite different.
-Each of $\|\Sigma\|$, $\|f_{\perp}^{\ast}\|_n^2$ and $\|f_{\perp}^{\ast}\|_P^2$ are random variables which depend on $\varphi_1,\ldots,\varphi_K$.
-We also do not have concentration of $\|f_{\perp}^{\ast}\|_n^2$ around $\|f_{\perp}^{\ast}\|_P^2$.
-
-\section{Kernel smoothing}
-Suppose we observe $(X_1,\wh{\theta}_1),\ldots,(X_n,\wh{\theta}_n)$, where $X_1,\ldots,X_n \in \Reals^d$ are sampled i.i.d from an unknown distribution $P$ on domain $\Omega$, and $\wh{\theta}_{1:n} = (\wh{\theta}_1,\ldots,\wh{\theta}_n) \in \Reals^n$.
-Our goal is to construct an estimator $\wh{f}$ such that the error $\|\wh{f} - f^{\ast}\|_P^2$ is as small as possible relative to $\|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n^2$.
-
-For a kernel function $\eta(\cdot): [0,\infty) \to (-\infty,+\infty)$, bandwidth $\varepsilon > 0$, and a distribution $Q$, the \emph{Nadaraya-Watson kernel smoother} $T_{\varepsilon,Q}$ is given by
-\begin{equation*}
-\bigl(T_{\varepsilon,Q}f)(x) := \frac{1}{d_Q(x)} \int_{\Omega} f(z)\eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dQ(z)
-\end{equation*}
-where $d_Q(x) = \int_{\Omega} \eta\bigl(\|z - x\|/\varepsilon\bigr) \,dQ(z)$. We denote $T_{\varepsilon,P_n}$ by $T_{\varepsilon,n}$, and $d_{P_n}$ by $d_n$.
-
-The estimated function $\wh{f}$ we will consider is a kernel smoother passed over $\wh{\theta}$,
-\begin{equation}
-\label{eqn:kernel_smoother_laplacian_eigenmaps}
-\wh{f}(x) := T_{\varepsilon,n}\wh{\theta}_{1:n}(x).
-\end{equation}
-In order to relate the out-of-sample error $\|\wh{f} - f^{\ast}\|_P$ to the in-sample error $\|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n$, we will assume the following.
-\begin{enumerate}[label=(A\arabic*)]
-	\item
-	\label{asmp:kernel}
-	The kernel function $\eta$ is compactly supported on a subset of $[0,1]$. Additionally, $\eta$ is Lipschitz on its support. Finally, for convenience we will assume that $\eta$ is normalized, 
-	\begin{equation*}
-	\int_{0}^{\infty} \eta(s) \,ds = 1.
-	\end{equation*}
-	\item 
-	\label{asmp:domain} There exists $\alpha > 0$ such that for every $x \in \mc{X}$, $d_{\nu}(x) \geq \alpha \cdot \varepsilon^d$.
-	\item 
-	\label{asmp:density}The distribution $P$ admits a density $p$ with respect to Lebesgue measure. The density $p \in C^1(\mc{X})$, and is bounded away from $0$ and $\infty$ for all $x \in \mc{X}$,
-	\begin{equation*}
-	0 < p_{\min} \leq p(x) < p_{\max} < \infty.
-	\end{equation*}
-\end{enumerate}
-Under Assumption~\ref{asmp:kernel}, we automatically have that $d_\nu(x) = \varepsilon^d$ for all $x \in \mc{X}_{\varepsilon}$.
-Assumption~\ref{asmp:domain} is thus a condition on the boundary $\partial \Omega$, and is implied for all sufficiently small $\varepsilon$ if the boundary is Lipschitz.
-Denoting $K(\cdot) = (\eta(\cdot))^2$, we note that Assumption~\ref{asmp:kernel} implies that $\eta$ is square-integrable, $\int_{0}^{\infty} K(s) \,ds < \infty$.
-\begin{lemma}
-	\label{lem:kernel_smoothing_laplacian_eigenmaps}
-	Suppose $f^{\ast} \in H^1(\mc{X})$ and $p \in C^1(\mc{X})$. Then with probability at least $1 - \delta - C\varepsilon^d\exp\{-Cn\varepsilon^d\}$, it holds that
-	\begin{equation*}
-	\|\wh{f} - f^{\ast}\|_P^2 \leq C\biggl(\|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n^2 + \frac{1}{\delta} \cdot \frac{\varepsilon^2}{n\varepsilon^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{\varepsilon,P}f^{\ast} - f^{\ast}\|_P^2\biggr)
-	\end{equation*}
-\end{lemma}
-Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data (on the order of $1/(n\varepsilon^d)$) by a factor of $\varepsilon^2$.
-The bias term, on the other hand, is standard (at least when $f^{\ast} \in C_0^s(\Omega)$).
-In particular the following estimate of the bias can be obtained.
-\begin{lemma}
-	\label{lem:kernel_smoothing_bias}
-	Suppose $f^{\ast} \in H_0^{s}(\Omega)$, $p \in C^{r}(\Omega)$ for $r = s - 1$, and that $\eta$ is an order-$s$ kernel. It follows that
-	\begin{equation*}
-	\|T_{\varepsilon,P}f^{\ast} - f^{\ast}\|_P^2 \leq C \varepsilon^{2s}.
-	\end{equation*}
-\end{lemma}
-\begin{proof}[Proof (when dimension $d = 1$).]
-	\textcolor{red}{(TODO)}
-\end{proof}
-
-Thus balancing bias and variance, using a smaller bandwidth ($\varepsilon \asymp n^{-1/(2r + d)}$) than is typical for kernel smoothing with noisy data, leads to an asymptotically negligible error ($n^{-2s/(2r + d)}$) compared to the minimax rate for estimating Sobolev functions.
-
-\begin{proof}[Proof of Lemma~\ref{lem:kernel_smoothing_laplacian_eigenmaps}]
-	By the triangle inequality, 
-	\begin{equation*}
-	\|\wh{f} - f^{\ast}\|_P \leq \|\wh{f} - T_{\varepsilon,n}f^{\ast}_{1:n}\|_P + \|T_{\varepsilon,n}f^{\ast}_{1:n} - f^{\ast}\|_P.
-	\end{equation*}
-	We analyze each of the two terms in the above expression separately. We will show that
-	\begin{equation}
-	\label{pf:kernel_smoothing_laplacian_eigenmaps_0}
-	\|\wh{f} - T_{\varepsilon,n}f^{\ast}_{1:n}\|_P \leq \frac{1}{\alpha}\bigl(1 + o_P(1)\bigr) \cdot \|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n
-	\end{equation}
-	and 
-	\begin{equation}
-	\label{pf:kernel_smoothing_laplacian_eigenmaps_0.5}
-	\|T_{\varepsilon,n}f^{\ast}_{1:n} - f^{\ast}\|_P^2 \leq \frac{C}{\delta}\frac{\varepsilon^2}{n\varepsilon^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{\varepsilon,P}f^{\ast} - f^{\ast}\|_P^2,
-	\end{equation}
-	the latter inequality holding with probability $1 - \delta$.
-
-	\textit{Proof of \eqref{pf:kernel_smoothing_laplacian_eigenmaps_0}.} We will begin by assuming that $\eta$ is positive, $\eta(z) \geq 0$ for all $z$, and return to prove the claim in its full generality at the end.
-
-	The left hand side of~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is equal to $\|T_{\varepsilon,n}(\wh{\theta}_{1:n} - f^{\ast}_{1:n})\|_P$, and~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is implied by showing that the operator $T_{\varepsilon,n}: L^2(X) \to L^2(\Omega)$ is (up to a constant) a contraction.
-We will show that for any $u \in L^2(X)$,
-	\begin{equation}
-	\label{pf:kernel_smoothing_laplacian_eigenmaps_1}
-	\|T_{\varepsilon,n}u\|_P^2 \leq \frac{1}{\alpha}\bigl(1 + o_P(1)\bigr) \|u\|_n^2.
-	\end{equation}
-	We begin with a pointwise estimate, which follows from Jensen's inequality:
-	\begin{equation*}
-	\Bigl(\bigl(T_{n,\varepsilon}u\bigr)(x)\Bigr)^2 \leq \frac{1}{d_n(x) n} \sum_{i = 1}^{n} \bigl(u(X_i)\bigr)^2 \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
-	\end{equation*}
-	(Note that it is here that we have used the positivity of $\eta(\cdot)$.) Integrating with respect to $P$ and exchanging sum with integral yields an upper bound on the $L^2(P)$ norm of $T_{n,\varepsilon}u$,
-	\begin{equation}
-	\label{pf:kernel_smoothing_laplacian_eigenmaps_2}
-	\|T_{n,\varepsilon}u\|_P^2 \leq \sum_{i = 1}^{n} u_i^2 \Bigl(\int_{\Omega} \frac{\eta(\|X_i - x\|/\varepsilon)}{n d_n(x)} p(x) \,dx \Bigr)
-	\end{equation}
-	The degree estimate $d_n(x)$ is quite close to $p(x)$.
-As we show in Lemma~\ref{lem:measure_of_ball_estimate}, if $\varepsilon = \omega((\log n/n)^{1/d})$ then the empirical measure $d_n(x)$ concentrates around $d_P(x)$ uniformly over $x \in \Omega$,
-	\begin{equation*}
-	\sup_{x \in \Omega}|d_n(x) - d_P(x)| = o_P(\varepsilon^d).
-	\end{equation*}
-	Moreover, it follows from the Lipschitz continuity of $p$ that for any $x \in \Omega$,
-	\begin{equation*}
-	d_P(x) = \int_{\Omega} \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) p(x') \,dx' = p(x) d_{\nu}(x) + O(\varepsilon^{d + 1}),
-	\end{equation*}
-	and thus by the triangle inequality $\sup_{x \in \Omega}|d_n(x) - d_\nu(x) p(x)| = o_P(\varepsilon^d)$.
-
-	As pointed out previously, if $x' \in \Omega_{\varepsilon}$ then $d_{\nu}(x') = \varepsilon^d$.
-On the other hand by assumption~\ref{asmp:domain} for any $x' \in \Omega, d_{\nu}(x') \geq \varepsilon^d/\alpha$.
-Thus
-	\begin{equation*}
-	\int_{\Omega} \frac{\eta(\|x' - x\|/\varepsilon)}{d_n(x')} p(x') \,dx' = \int_{\Omega} \frac{\eta(\|x' - x\|/\varepsilon)}{d_{\nu}(x') - o_P(\varepsilon^d)} \,dx' \leq 
-	\begin{dcases*}
-	\frac{1}{1 - o_P(1)}, & \textrm{if $x \in \Omega_{2\varepsilon}$} \\
-	\frac{1}{\alpha} \cdot \frac{1}{1 - o_P(1)}, & \textrm{otherwise.}
-	\end{dcases*}
-	\end{equation*}
-	Plugging back in to~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_2}, we deduce that
-	\begin{equation*}
-	\|T_{n,\varepsilon}u\|_P^2 \leq \sum_{i = 1}^{n} u_i^2 \biggl(\1\Bigl\{x \in \Omega_{2\varepsilon}\Bigr\} \frac{1}{1 - o_P(1)} + \1\Bigl\{x \in (\partial\Omega)_{2\varepsilon}\Bigr\} \frac{1}{\alpha} \frac{1}{1 - o_P(1)}\biggr).
-	\end{equation*}
-	This establishes~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_1}, and in turn~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0}.
-
-	Now, suppose $\eta(\cdot)$ can take on negative values.
-In this case, we have $\eta(\cdot) = \mathrm{sgn}(\eta(\cdot))K^{1/2}(\cdot)$.
-Let $d_n(x;K^{1/2})$ be the empirical degree with respect to kernel function $K^{1/2}$.
-Then
-	\begin{align*}
-	T_{n,\varepsilon}(u)(x) & = \frac{1}{d_n(x)n} \sum_{i = 1}^{n} \bigl(u(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr) \\
-	& = \frac{d_n(x;K^{1/2})}{d_n(x)}\frac{1}{d_n(x;K^{1/2})n} \sum_{i = 1}^{n} \bigl(u(X_i)\bigr) \mathrm{sgn}\biggl(\eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr)\biggr) K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
-	\end{align*}
-	Taking the square and applying Jensen's inequality, we deduce that
-	\begin{align*}
-	\bigl(T_{n,\varepsilon}u(x)\bigr)^2 & \leq \biggl(\frac{d_n(x;K^{1/2})}{d_n(x)}\biggr)^2 \cdot \frac{1}{d_n(x;K^{1/2})n}\sum_{i = 1}^{n} \bigl(u(X_i)\bigr)^2 K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr) \\
-	& \leq \frac{(1 + o_P(1))\|\eta\|_{\infty}^2}{\alpha^2} \cdot \frac{1}{d_n(x;K^{1/2})n}\sum_{i = 1}^{n} \bigl(u(X_i)\bigr)^2 K^{1/2}\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr),
-	\end{align*}
-	with the latter inequality following by Lemma~\ref{lem:measure_of_ball_estimate} and assumption~\ref{asmp:kernel}.
-From here the proof proceeds along similar lines to the case where $\eta$ is non-negative.
-
-	\textit{Proof of~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0.5}.}  
-	We will begin by introducing some notation. For a function $u \in L^2(\Omega)$, let
-	\begin{equation*}
-	L_{\varepsilon,n}u(x) := T_{\varepsilon,n}u(x) - u(x) = \frac{1}{d_n(x)n} \sum_{i = 1}^{n} \bigl(u(x) - u(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr)
-	\end{equation*}
-	and let
-	\begin{equation*}
-	\wt{L}_{\varepsilon,n}u(x) := \frac{d_n(x)}{d_P(x)}L_{\varepsilon,n}u(x) = \frac{1}{d_P(x)n} \sum_{i = 1}^{n} \bigl(u(x) - u(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
-	\end{equation*}
-	We have already shown that $d_n(x)/d_P(x) = (1 + o_P(1))$, and so we will focus our analysis on upper bounding $\Ebb[\|\wt{L}_{\varepsilon,n}f^{\ast}\|_P^2]$.
-We begin by providing expressions for the pointwise mean and variance of $\wt{L}_{\varepsilon,n}f^{\ast}(x)$:
-	\begin{equation*}
-	\Ebb\bigl[\wt{L}_{\varepsilon,n}f^{\ast}(x)\bigr] = T_{P,\varepsilon}f^{\ast}(x) - f^{\ast}(x)
-	\end{equation*}
-	and
-	\begin{align*}
-	\mathbb{V}\bigl[\wt{L}_{\varepsilon,n}f^{\ast}(x)\bigr] & = \frac{1}{(d_P(x))^2 n} \mathbb{V}\bigl[\bigl(f^{\ast}(X) - f^{\ast}(x)\bigr) \eta\bigl(\|X - x\|/\varepsilon\bigr)\bigr] \\
-	& \leq \frac{1}{p_{\min}^2 \alpha^2 \varepsilon^{2d}} \mathbb{E}\bigl[\bigl(f^{\ast}(X) - f^{\ast}(x)\bigr)^2 K\bigl(\|X - x\|/\varepsilon\bigr)\bigr] \\
-	& = \frac{1}{p_{\min}^2 \alpha^2 \varepsilon^{2d}} \int_{\Omega} \bigl(f^{\ast}(x') - f^{\ast}(x)\bigr)^2 K(\|x' - x\|/\varepsilon) p(x) \,dx.
-	\end{align*}
-	Then, integrating with respect to $P$ over $\Omega$, we obtain
-	\begin{align*}
-	\Ebb\bigl[\|\wt{L}_{\varepsilon,n}f^{\ast}\|_P^2\bigr]  & \leq \|T_{P,\varepsilon}f^{\star} - f^{\star}\|_{P}^2 + \frac{1}{p_{\min}^2 \alpha^2 \varepsilon^{2d}} E_{\varepsilon}(f^{\ast}) \\
-	& \leq \|T_{P,\varepsilon}f^{\star} - f^{\star}\|_{P}^2 + \frac{C\varepsilon^2}{p_{\min}^2 \alpha^2 \varepsilon^{d}} |f|_{H^1(\Omega)}^2
-	\end{align*}
-	where $E_{\varepsilon}(f^{\ast}) := \int_{\Omega} \int_{\Omega} (f^{\ast}(x') - f^{\ast}(x))^2 K(\|x' - x\|/\varepsilon) p(x') p(x) ,\dx' \,dx$, and we have shown in our AISTATS paper that $E_{\varepsilon}(f^{\ast}) \leq C \varepsilon^{d + 2} |f|_{H^1(\Omega)}^2$, justifying the second inequality.
-Consequently,
-	\begin{align*}
-	\|T_{n,\varepsilon}f^{\ast}_{1:n} - f^{\ast}\|_P^2 \leq (1 + o_P(1))^2 \|\wt{L}_{\varepsilon,n}f^{\ast}\|_P^2 & \leq \frac{\bigl(1 + o_P(1)\bigr)^2}{\delta} \mathbb{E}\bigl[\|\wt{L}_{\varepsilon,n}f^{\ast}\|_P^2\bigr] \\
-	& \leq \frac{\bigl(1 + o_P(1)\bigr)^2}{\delta}\biggl(\|T_{P,\varepsilon}f^{\star} - f^{\star}\|_{P}^2 + \frac{C\varepsilon^2}{p_{\min}^2 \alpha^2 \varepsilon^{d}} |f|_{H^1(\Omega)}^2\biggr)
-	\end{align*}
-	with probability at least $1 - \delta$, establishing~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0.5}.
-\end{proof}
-
-Recall the notation $d_P(x) = \int \eta(\|z - x\|/\varepsilon) \,dP(z)$, $d_n(x) = d_{P_n}(x)$. Recall also that $K(\cdot) = |\eta(\cdot)|^2$. 
-\begin{lemma}
-	\label{lem:measure_of_ball_estimate}
-	Let $\eta$ be a kernel compactly supported and Lipschitz on $[0,1]$. There exists a constant $C$ that does not depend on $\varepsilon$ or $\eta$, such that
-	\begin{equation*}
-	\sup_{x \in \Omega}|d_n(x) - d_P(x)| \leq C \sqrt{\frac{\varepsilon^d}{n}},
-	\end{equation*}
-	and
-	\begin{equation*}
-	\sup_{x \in \Omega}|d_n(x;K^{1/2}) - d_P(x;K^{1/2})| \leq C \sqrt{\frac{\varepsilon^d}{n}},
-	\end{equation*}
-	with probability at least $1 - \textcolor{red}{(?)}\exp\{-Cn\varepsilon^d\}$. 
-\end{lemma}
-
-\begin{proof}
-	Since $\eta$ is compactly supported on $[0,1]$, a standard computation implies that since for any $x \in \mc{X}$,
-	\begin{equation*}
-	\Var\bigl[d_n(x)\bigl] = \frac{1}{n} \Var[\eta(\|X - x\|/\varepsilon)] \leq \frac{1}{n}\|\eta\|_{\infty}^2 \nu_d \varepsilon^d,
-	\end{equation*}
-	where we note that $\|\eta\|_{\infty} < \infty$ follows from the assumption that $\eta$ is Lipschitz. It follows from Bernstein's inequality that 
-	\begin{equation*}
-	\Pbb\biggl(|d_n(x) - d_P(x)| \leq C \sqrt{\frac{\varepsilon^d}{n}} \biggr) \leq 2\exp\biggl\{-\frac{t^2 \varepsilon^d}{C(\varepsilon^d + \delta \sqrt{\varepsilon^d/n})}\biggr\}.
-	\end{equation*}
-	for $C = \|\eta\|_{\infty} \nu_d$. \textcolor{red}{(...)}
-\end{proof}
-
-\section{Nystr\"{o}m extension of Laplacian eigenvectors}
-The Nystr\"{o}m extension $\varphi_k$ of an eigenvector $v_k$ of the graph Laplacian $L$ is
-\begin{equation}
-\label{eqn:nystrom}
-\varphi_k(x) := \frac{1}{d_n(x) - \lambda_k} \sum_{i = 1}^{n} v_{k,i} K\Bigl(\frac{\|X_i - x\|}{r}\Bigr)
-\end{equation}
-where $d_n(x) := \sum_{i = 1}^{n} K(\|X_i - x\|/r)$, and we assume $d_n(x) > \lambda_k$ for all $x \in \Omega$ and $k = 1,\ldots,K$.
-There are certain nice properties of the Nystr\"{o}m extension.
-It is an interpolant---that is, for any $\theta \in \mathrm{span}\{v_1,\ldots,v_K\}$ the function $f \in \Leb^2(\Omega)$ will agree with $\theta$ at sample design points---and it requires no information on $\Omega$ to be well-defined.
-However it is also tricky to analyze, because it involves taking $\theta \in \Leb^2(X)$, transforming it into eigenspace, and then applying a \emph{different} smoother to each eigenvector.
-To build some intuition, we will warm up by analyzing piecewise constant extension, which is much easier to tackle theoretically, if suboptimal for $f^{\ast} \in H^s(\mc{X})$ when $s > 1$.
-
-\subsection{Piecewise constant extension}
-Both $\|\Sigma\|$ and $\|f_{\perp}^{\ast}\|_P^2$ are easy to analyze when $\varphi_1,\ldots,\varphi_k$ are taken to be piecewise constant extensions of eigenvectors $v_k$ of the Laplacian matrix $L$.
-Let $\{V_1,\ldots,V_n\}$ partition the domain $\Omega$ into $n$ cells, each of probability mass $\Pbb(V_j) = 1/n$, such that $x_j \in V_j$.
-The piecewise constant extension operator $P^{\star}: L^2(X)\to L^2(\Omega)$ is
-\begin{equation}
-\label{eqn:piecewise_extension}
-\bigl(P^{\star}\theta\bigr)(x) = \sum_{j = 1}^{n} \theta_j \1\{x \in V_j\},
-\end{equation}
-and we take $\varphi_k = P^{\star}v_k$ for $k = 1,\ldots,K$.
-
-It is clear that $\dotp{{\varphi}_k}{{\varphi}_{\ell}}_{P} = \dotp{v_k}{v_{\ell}}_{n} = \delta_{ij}$, so that $\Sigma = I$ and $\|\Sigma\| = 1$.
-Moreover, we have that $f_K^{\ast} = P^{\ast}\wt{f}_K$, so that
-\begin{equation*}
-\|P^{\star}f_{1:n}^{\ast} - {f}_K^{\ast}\|_P = \|f^{\ast} - \wt{f}_K\|_n = \|f^{\ast} - f_K^{\ast}\|_n;
-\end{equation*}
-the term $\|f^{\ast} - \wt{f}_K\|_n$ is an in-sample bias term that we have already analyzed.
-It remains to upper bound $\|P^{\star}f^{\ast} - f^{\ast}\|_P$, which we can do using crude but sufficient techniques when $f^{\ast} \in C^1(\Omega)$.
-
-\subsection{Nystr\"{o}m extension}
-
-The analysis of the Nystr\"{o}m extension seems challenging.
-At present I have suitable bounds on zero out of three of the functionals in question ($\|\Sigma\|$, $\|f_{\perp}^{\ast}\|_n^2$ and $\|f_{\perp}^{\ast}\|_P^2$).
-
-\paragraph{Energy of population-level covariance $(\|\Sigma\|)$.}
-The following representation of $\|\Sigma\|$ seems useful:
-\begin{equation}
-\label{eqn:covariance_energy}
-\|\Sigma\| = \sup_{\substack{f \in \mathrm{span}\{\varphi_1,\ldots,\varphi_K\} \\ \|f\|_n = 1}} \|f\|_P.
-\end{equation}
-The equivalence~\eqref{eqn:covariance_energy} suggests a useful intermediary step in upper bound $\|\Sigma\|$ is an upper bound on $\|\varphi_k\|_P^2$.
-To do so, it is useful express the Nystrom extension of $v_k$ in terms of a kernel smoother.
-Recall that $T_{r,n}$ is the kernel smoother,
-\begin{equation*}
-\bigl(T_{r,n}\theta\bigr)(x) := \frac{1}{d_n(x)} \sum_{i = 1}^{n} \theta_i K\biggl(\frac{\|X_i - x\|}{r}\biggr).
-\end{equation*}
-Then constructing $\varphi_k$ is equivalent to first passing a kernel smoother over $v_k$, then applying a multiplication operator, i.e
-\begin{equation}
-\label{eqn:nystrom_extension_kernel_smoother_representation}
-\varphi_k(x) = \frac{d_n(x)}{d_n(x) - \lambda_k} \bigl(T_{r,n}v_k\bigr)(x).
-\end{equation}  
-The operator $T_{r,n}$ is a contraction, in the sense that $\|T_{r,n}\theta\|_P \leq C \|\theta\|_n$, as the following derivations make clear:
-\begin{align*}
-\|T_{r,n}\theta\|_P^2 & = \int \Biggl(\frac{1}{d_n(x)} \sum_{i = 1}^{n} \theta_i K\biggl(\frac{\|X_i - x\|}{r}\biggr)\Biggr)^2 \,dP(x) \\
-& \leq \sum_{i = 1}^{n} \theta_i^2 \int \frac{1}{d_n(x)}  K\biggl(\frac{\|X_i - x\|}{r}\biggr) \,dP(x) \\
-& \leq \|\theta\|_n^2 \cdot \frac{d_{\max}(P)}{d_{\min}(P_n)}.
-\end{align*}
-(One can do better than this, actually, but this suffices for our purposes).
-
-Therefore from~\eqref{eqn:nystrom_extension_kernel_smoother_representation}, and Holder's inequality, we have that
-\begin{equation}
-\label{eqn:nystrom_extension_l2p_norm}
-\|\varphi_k\|_P \leq \sup_{x \in \Omega} \biggl\{\frac{d_n(x)}{d_n(x) - \lambda_k} \biggr\} \cdot  \|T_rv_k\|_P \leq \frac{d_{\min}(P_n)}{d_{\min}(P_n) - \lambda_k/n} \cdot \sqrt{ \frac{d_{\max}(P)}{d_{\min}(P_n)}}.
-\end{equation}
-For \textcolor{red}{typical} choices of $K$ and $r$, we have that $d_{\min}(P_n) \gg \lambda_K/n$, and therefore the upper bound~\eqref{eqn:nystrom_extension_l2p_norm} is $O(1)$, which suffices for our purposes.
-
-Unfortunately, notwithstanding my prior claim, I don't yet see the right way to leverage this into a bound on $\|\Sigma\|$.
-For a given $\theta = \sum_{k = 1}^{K} \beta_k v_k$, and the resulting extrapolant $f = \sum_{k = 1}^{K} \beta_k \varphi_k(x)$, we can write $f$ in terms of $T_r\theta$ and a remainder term as follows,
-\begin{equation*}
-f(x) = \bigl(T_r\theta\bigr)(x) + \sum_{k = 1}^{K} \beta_k \frac{\lambda_k}{d_n(x) - \lambda_k} \bigl(T_rv_k\bigr)(x).
-\end{equation*}
-However, my best analysis of the remainder term, which I state here, is doomed by accumulation of error:
-\begin{align*}
-\biggl\|\sum_{k = 1}^{K} \beta_k \frac{\lambda_k}{d_n(x) - \lambda_k} \bigl(T_rv_k\bigr)(x)\biggr\|_P^2 & \overset{(i)}{\leq} K \cdot \sum_{k = 1}^{K} \beta_k^2  \biggl(\frac{\lambda_k/n}{d_{\min}(P_n) - \lambda_k/n}\biggr)^2 \cdot  \|T_rv_k\|_P^2 \\
-& \overset{(ii)}{\leq} K \biggl(\frac{\lambda_K/n}{d_{\min}(P_n) - \lambda_K/n}\biggr)^2 \cdot \frac{d_{\max}(P)}{d_{\min}(P_n)} \\
-& \asymp \frac{K \lambda_K^2}{n^2d_{\min}(P_n)}.
-\end{align*}
-To see that this bound is insufficient: recall the optimal choice $K_{\star} \asymp n^{d/(d + 2s)}$, that the eigenvalue $\lambda_K \asymp nr^{d + 2}K^{2/d}$ and that $d_{\min}(P_n) \asymp r^d$.
-Thus the bound reduces to $n^{\gamma}$ with exponent $\gamma = 3d/(2s + d) - 4/d$, which is greater than $0$ for most values of $d$.
-Moreover, the looseness in inequality $(ii)$ did not kill us---replacing $\lambda_K$ by $\lambda_2$ on the right hand side of $(ii)$ still results in a bound growing with $n$.
-On the other hand, the inequality $(i)$ was rather gross---it is tight only when $\varphi_k \propto \varphi_{\ell}$ for all $k,\ell = 1,\ldots,K$, and it seems likely that we can do better.
-
-\paragraph{Distance between $f^{\ast}$ and subspace spanned by $\varphi_1,\ldots,\varphi_K$ $(f_{\perp}^{\ast})$.}
-
-It seems like a useful starting place would be to take $f^{\ast} \propto \psi_k$, where $\psi_k$ is the $k$th eigenvector of the continuum weighted Laplace-Beltrami operator, and show that this is close to $\mathrm{span}\{\varphi_1,\ldots,\varphi_K\}$.
-I do not know an upper bound on the distance between $\psi_k$ and $\mathrm{span}\{\varphi_1,\ldots,\varphi_K\}$, but I do know such a bound for a different extension of $v_1,\ldots,v_K$.
-
-\paragraph{$(\|f_{\perp}^{\ast}\|_n)$.}
-
-This one I frankly have no thoughts on how to analyze. I don't even understand it all that well, as evidenced by the fact that I can't name it.
-
-\clearpage
-
-\section{Prediction error for least-squares estimators: random design, fixed basis.}
-In this section, we will assume that $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported over $\Omega \subseteq \Rd$, and that $\phi_1,\ldots,\phi_k$ are deterministic functions which in particular do not depend on $X_1,\ldots,X_n$.
-
-The prediction error of $\wh{f}$ is $\mathbb{E}_{(X,Y)}[(Y - \wh{f}(X))^2]$, where $X \sim P$ and $\mathbb{E}[Y|X] = f^{\ast}(X)$ are independent of $X_1,\ldots,X_n$, and we additionally assume $Y - \mathbb{E}[Y|X]$ is independent of $X$ (i.e.
-homoskedastic noise).
-The prediction error can be decomposed into the sum of the $L^2(P)$ error and an irreducible error,
-\begin{equation}
-\mathbb{E}_{(X,Y)}[(Y - \wh{f}(X))^2] = \|\wh{f} - f^{\ast}\|_P^2 + \mathbb{E}_{(X,Y)}[(Y - f^{\ast}(X))^2],
-\end{equation}
-justifying our interest in the $\|\cdot\|_P^2$ error.
-
-We will apply Lemma~\ref{lem:l2p_error_fixed_design}, and then use our assumptions on the design points and basis functions to control the right hand side of~\eqref{eqn:l2p_error_fixed_design}.
-In particular, we have the following.
-\begin{lemma}
-	content...
-\end{lemma}
-
-\section{Miscellaneous}
-I put things here which do not belong in any other section.
-
-\subsection{Proof of Lemma~\ref{lem:kernel_smoothing_bias}, $d = 1$ and $p$ uniform}
-\label{subsec:kernel_smoothing_bias_pf_1d_uniform_p}
-Without loss of generality take the domain $\Omega = (0,1)$. Recall that we are interested in upper bounding
-\begin{equation*}
-\|T_{P,\varepsilon}u - u\|_{P}^2 = \int_{\Omega} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x) \leq C\varepsilon^{2k}
-\end{equation*}
-where $L_{P,\varepsilon}$ is a non-local approximation to the continuum weighted Laplacian, given by
-\begin{equation*}
-L_{P,\varepsilon}u(x) := \frac{1}{d_P(x)}\int_{\Omega} \bigl(u(z) - u(x)\bigr) \eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dP(z).
-\end{equation*}
-Let $\Omega_{\varepsilon} := \{x: B(x,\varepsilon) \in \Omega\}$ and let $\partial_{\varepsilon}\Omega := \Omega \setminus \Omega_{\varepsilon}$. Clearly
-\begin{equation*}
-\|L_{P,\varepsilon}u\|_{P}^2 = \underbrace{\int_{\Omega_{\varepsilon}} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_1} + \underbrace{\int_{\partial_{\varepsilon}\Omega} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_2}
-\end{equation*}
-and we now separately upper bound $I_1$ and $I_2$.
-
-\paragraph{Upper bound on $I_1$.}
-We begin with a pointwise analysis of $L_{p,\varepsilon}u(x)$ at a given $x \in \Omega_{\varepsilon}$.
-We may assume without loss of generality that $u \in C_0^{\infty}(\Omega)$, since $C_0^{\infty}(\Omega)$ is dense in $H_0^k(\Omega)$ and $\|T_{P,\varepsilon}u - u\|_P^2$ is continuous with respect to $\|\cdot\|_{H_0^k}$.
-Since $u$ is smooth, for every $x,z \in \Omega$, we have the $k$th-order Taylor expansion
-\begin{equation*}
-u(z) - u(x) = \sum_{j = 1}^{k - 1} (z - x)^{j} u^{(j)}(x)  + \frac{(z - x)^k}{(k - 1)!}\int_{0}^{1}(1 - t)^{k - 1}u^{(k)}(x + t(z - x))\,dt.
-\end{equation*}
-As a result,
-\begin{equation}
-\label{pf:kernel_smoothing_bias_1d_1}
-\begin{aligned}
-L_{P,\varepsilon}u(x) = \frac{1}{d_P(x)} & \Biggl\{\sum_{j = 1}^{k - 1} u^{(j)}(x) \int_{\Omega}(z - x)^{j}\eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr)\,dz + \\ 
-& \frac{1}{(k - 1)!} \int_{\Omega} \int_{0}^{1} (1 - t)^k (z - x)^k u^{(k)}(x + t(z - x)) \eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dt \,dz \Biggr\}
-\end{aligned}
-\end{equation}
-
-Now, since $x \in \Omega_{\varepsilon}$ then all $z \in B(x,\varepsilon)$ also belong to $\Omega$. For one thing, this implies
-\begin{equation*}
-d_P(x) = \int_{\Omega} \eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dz = \varepsilon \cdot \int_{B(0,1)}  \eta(v) \,dv = \epsilon.
-\end{equation*}
-Additionally, we have that all $j = 1,\ldots,k - 1$, 
-\begin{equation*}
-\int_{\Omega}(z - x)^{k}\eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dz = \varepsilon^j \cdot \int_{B(0,1)} v^j \eta(v) \,dv = 0,
-\end{equation*}
-where the second equality follows because $\eta$ is a $k$th-order kernel.
-Consequently $L_{P,\varepsilon}$ depends only on the $k$th derivative of $u$: changing variables to $v = (z - x)/\varepsilon$, we have
-\begin{align*}
-L_{P,\varepsilon}u(x) & = \frac{1}{\varepsilon} \cdot \frac{1}{(k - 1)!} \int_{0}^{1} \int_{\Omega} (1 - t)^k (z - x)^k u^{(k)}(x + t(z - x)) \eta\biggl(\frac{\|z - x\|}{\varepsilon}\biggr) \,dz \,dv \\
-& = \frac{\varepsilon^{k}}{(k - 1)!} \int_{0}^{1} \int_{B(0,1)} (1 - t)^k v^k u^{(k)}(x + t\varepsilon v) \eta(|v|) \,dv \,dt
-\end{align*} 
-and as a result
-\begin{equation*}
-\Bigl(L_{P,\varepsilon}u(x)\Bigr)^2 = \frac{\varepsilon^{2k}}{(k!)^2} {\sigma_\eta}^2 \bigl(u^{(k)}(x)\bigr)^2 + \frac{\varepsilon^{2k}}{((k - 1)!)^2}A_1(x)A_{2}(x),
-\end{equation*}
-where $\sigma_\eta = \int v^k \eta(|v|) \,dv$ and 
-\begin{align*}
-A_1(x) & := \int_{0}^{1} \int_{B(0,1)} (1 - t)^k v^k \bigl(u^{(k)}(x + t\varepsilon v) - u^{(k)}(x)\bigr) \eta(|v|) \,dv \,dt \\
-A_2(x) & := \int_{0}^{1} \int_{B(0,1)} (1 - t)^k v^k \bigl(u^{(k)}(x + t\varepsilon v) + u^{(k)}(x)\bigr) \eta(|v|) \,dv \,dt.
-\end{align*}
-Integrating over all $x \in \Omega_{\varepsilon}$, we have that
-\begin{align*}
-I_1 = \int_{\Omega_{\varepsilon}} \bigl(L_{P,\varepsilon}u(x)\bigr)^2 \,dx & = \frac{\varepsilon^{2k}}{(k!)^2} \sigma_\eta^2 \int_{\Omega_{\varepsilon}} \bigl(u^{(k)}(x)\bigr)^2 \,dx + \frac{\varepsilon^{2k}}{((k - 1)!)^2} \int_{\Omega_{\varepsilon}}A_1(x) A_2(x) \,dx \\
-& \leq \frac{\varepsilon^{2k}}{(k!)^2} \Bigl(\sigma_\eta^2 \|u^{(k)}\|_{P}^2 + k^2 \|A_1\|_P \|A_2\|_P\Bigr).
-\end{align*}
-It remains to show that $\|A_1\|_P\|A_2\|_P = o(1)$.
-
-\textcolor{red}{(TODO): Follow through on this calculation, referring to pages 192 - 195 of Tsybakov's book.}
-
-\paragraph{Upper bound on $I_2$.}
-We will prove the upper bound
-\begin{equation}
-\label{pf:kernel_smoothing_bias_1d_2}
-\int_{0}^{\varepsilon} \bigl|L_{P,\varepsilon}u(x)\bigr|^2 \,dx \leq C \varepsilon^{2k + 1}\|u^{(k)}\|_P^2,
-\end{equation}
-with the bound on $\int_{1 - \varepsilon}^{1} \bigl|L_{P,\varepsilon}u\bigr|^2$ following in a completely analogous fashion.
-
-Unlike the previous case, when $x$ is near the boundary of $\Omega$ we cannot use the higher-order property of $\eta$ to cancel the order-$1$ up through order-$(k - 1)$ derivatives of $u$.
-Instead, we will use the zero-trace property to show that $u$ itself must be quite small.
-Taking a Taylor expansion of $u$ around $0$, we have that for any $x \in \Omega$,
-\begin{align}
-u(x) & = u(0) + \sum_{j = 1}^{k - 1} x^j u^{(j)}(0) + \frac{1}{(k - 1)!} \int_{0}^{x}u^{(k)}(s)(x - s)^{k - 1}\,ds \nonumber \\
-& = \frac{1}{(k - 1)!} \int_{0}^{x}u^{(k)}(s)(x - s)^{k - 1}\,ds. \label{pf:kernel_smoothing_bias_1d_3}
-\end{align}
-with the second inequality following since $u \in C_c^{\infty}(\Omega)$ implies $u(0) = 0$ as well as $u^{(j)}(0) = 0$ for $j = 1,\ldots,k - 1$.
-To make use of this upper bound on $u(x)$, we upper bound $|L_{P,\varepsilon}u(x)|^2$ using the Cauchy-Schwarz inequality,
-\begin{align*}
-|L_{P,\varepsilon}u(x)|^2 & = \Biggl(\frac{1}{d_P(x)} \int_{\Omega} \bigl(u(x) - u(z)\bigr) \eta\biggl(\frac{|x - z|}{\varepsilon}\biggr) \,dz\Biggr)^2 \\
-& \leq \frac{1}{d_P(x)}\biggl(\int_{\Omega} K^{1/2}\biggl(\frac{x - z}{\varepsilon}\biggr) \,dz\biggr) \cdot \frac{1}{d_P(x)}\int_{\Omega} \bigl(u(x) - u(z)\bigr)^2 K^{1/2}\biggl(\frac{|x - z|}{\varepsilon}\biggr) \,dz \\
-& \leq 2\frac{\|\eta\|_{\infty}}{\alpha^2\varepsilon} \cdot \int_{\Omega} \bigl(u(x)^2 + u(z)^2\bigr) K^{1/2}\biggl(\frac{|x - z|}{\varepsilon}\biggr) \,dz,
-\end{align*}
-where in the final inequality we have used the estimates $d_P(x;K^{1/2}) \leq \|\eta\|_{\infty} \varepsilon$ and $d_P(x) \geq (\varepsilon\alpha)$.
-
-We use these pointwise estimates to upper bound~\eqref{pf:kernel_smoothing_bias_1d_2}.
-Since $\eta$ is compactly supported, if $x \in (0,\varepsilon]$ and $\eta(|x - z|/\varepsilon) > 0$ then $z \in (0,2\varepsilon]$.
-As a result,
-\begin{align*}
-\int_{0}^{\varepsilon} |L_{P,\varepsilon}u(x)|^2 \,dP(x) & \leq 2\frac{\|\eta\|_{\infty}}{\alpha^2\varepsilon} \int_{0}^{\varepsilon}  \int_{\Omega} \bigl(u(x)^2 + u(z)^2\bigr) K^{1/2}\biggl(\frac{|x - z|}{\varepsilon}\biggr) \,dz \,dx \\
-& \leq 2\frac{\|\eta\|_{\infty}}{\alpha^2\varepsilon} \int_{0}^{2\varepsilon}  \int_{0}^{2\varepsilon} \bigl(u(x)^2 + u(z)^2\bigr) K^{1/2}\biggl(\frac{|x - z|}{\varepsilon}\biggr) \,dz \,dx \\
-& \leq 4\frac{\|\eta\|_{\infty}^2}{\alpha^2} \int_{0}^{2\varepsilon} (u(x))^2 \,dx,
-\end{align*}
-where the last inequality follows by Fubini's Theorem.
-
-It remains to upper bound the $\Leb^2$ norm of $u$ over $(0,2\varepsilon]$. Here we use the bound~\eqref{pf:kernel_smoothing_bias_1d_3} and the Cauchy-Schwarz inequality to conclude that 
-\begin{equation*}
-\int_{0}^{2\varepsilon} \bigl(u(x)\bigr)^2 \,dx \leq \int_{0}^{2\varepsilon} \biggl(\int_{0}^{x} u^{(k)}(s)(x - s)^{k - 1} \,ds\biggr)^2 \,dx \leq \varepsilon^{2k - 1} \int_{0}^{2\varepsilon} \int_{0}^{x} \bigl(u^{(k)}(s)\bigr)^2 \,ds \,dx.
-\end{equation*}
-Finally, switching the order of integration and substituting $v = s/(2\varepsilon)$ gives
-\begin{align*}
-\int_{0}^{2\varepsilon} \int_{0}^{x} \bigl(u^{(k)}(s)\bigr)^2 \,ds \,dx = \int_{0}^{2\varepsilon} \int_{x}^{2\varepsilon} \bigl(u^{(k)}(s)\bigr)^2 \,dx \,ds \leq 2\varepsilon \int_{0}^{2\varepsilon} \bigl(u^{(k)}(s)\bigr)^2 \,ds = 4\varepsilon^2 \|u^{(k)}\|_P^2,
-\end{align*}
-establishing~\eqref{pf:kernel_smoothing_bias_1d_2} upon choosing $C = \frac{16\|\eta\|_{\infty}^2}{\alpha^2}$ in the statement of the inequality.
-
-\subsection{Proof of Lemma~\ref{lem:kernel_smoothing_bias}, $d = 1$ and $p \in C^k$.}
-\label{subsec:kernel_smoothing_bias_pf_1d_smooth_p}
-As in~\ref{subsec:kernel_smoothing_bias_pf_1d_uniform_p}, we take $\Omega = (0,1)$, let
-\begin{equation*}
-\|L_{P,\varepsilon}u\|_{P}^2 = \underbrace{\int_{\Omega_{\varepsilon}} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_1} + \underbrace{\int_{\partial_{\varepsilon}\Omega} \bigl(L_{\varepsilon,P}u(x)\bigr)^2 \,dP(x)}_{:=I_2},
-\end{equation*}
-and turn to separately upper bounding $I_1$ and $I_2$.
-
-\paragraph{Upper bound on $I_1$.}
-We will show that for all $x \in \Omega_{\varepsilon}$,
-\begin{equation*}
-L_{P,\varepsilon}u(x) = \varepsilon^{k} \frac{\sigma_\eta}{k!}\Bigl( \frac{d^k}{dx^k}(up)(x) - p^{(k)}(x)u(x) + r(x)\Bigr)
-\end{equation*}
-for a remainder term $r$ satisfying $\int_{\Omega_{\varepsilon}} \bigl(r(x)\bigr)^2 p(x) \,dx  = o(1)$ as $\varepsilon \to 0$. This in turn will imply
-\begin{align*}
-I_1 = \int_{\Omega_{\varepsilon}} |L_{P,\varepsilon}u(x)|^2 \,dx & = \varepsilon^{2k} \frac{\sigma_\eta^2}{(k!)^2}\Bigl(\int_{\Omega_{\varepsilon}} \Bigl( \frac{d^k}{dx^k}(up)(x) - p^{(k)}(x)u(x) + r(x)\Bigr)^2  p(x) \,dx \\
-& \leq 3 \varepsilon^{2k} \frac{\sigma_\eta^2}{(k!)^2} \bigl( |up|_{H^k(P)}^2 + |p^{(k)}u|_{\Leb^2(P)}^2 + o(1)\bigr) \\
-& \leq C \varepsilon^{2k} \frac{\sigma_\eta^2}{(k!)^2} \Bigl( \|p\|_{C^k(P)}^2\bigl(|u|_{H^k(P)}^2 + |u|_{\Leb^2(P)}^2\bigr) + o(1) \Bigr),
-\end{align*}
-with the last line following from Young's inequality.
-
-It remains to show the pointwise estimate \textcolor{red}{(...)}.
+\section{Proof of~\eqref{eqn:local_weyl_law}}
 

### 2021-12-22 notes/out_of_sample_extension.tex
Some miscellaneous notes.

--- 2021-09-04 manuscripts/laplacian_eigenmaps/annals_of_statistics/supplement.tex
@@
-\documentclass[aos]{imsart}
+\documentclass{article}
 
-%% Packages
-\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
-% \RequirePackage[numbers]{natbib}
-\RequirePackage[authoryear]{natbib}
-\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
-\RequirePackage{graphicx}
-\usepackage{mathtools, enumitem}
+%%% Begin Ryan's template
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{times}          % times font
+
+\usepackage[round]{natbib}
+\usepackage{amssymb,amsmath,amsthm,bbm}
+\usepackage[margin=1in]{geometry}
+\usepackage{verbatim,float,url,dsfont}
+\usepackage{graphicx,subfigure,psfrag}
+\usepackage{algorithm,algorithmic}
+\usepackage{mathtools,enumitem}
+\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}
 \usepackage{multirow}
 
-\startlocaldefs
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% For Axiom, Claim, Corollary, Hypothesis, %%
-%% Lemma, Theorem, Proposition              %%
-%% use \theoremstyle{plain}                 %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\theoremstyle{plain}
+% Theorems and such
 \newtheorem{theorem}{Theorem}
@@
 \theoremstyle{definition}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% For Assumption, Definition, Example,     %%
-%% Notation, Property, Remark, Fact         %%
-%% use \theoremstyle{remark}                %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\theoremstyle{remark}
 \newtheorem{remark}{Remark}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Please put your definitions here:        %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%% Begin Ryan's definitions
+% Assumption
+\newtheorem*{assumption*}{\assumptionnumber}
+\providecommand{\assumptionnumber}{}
+\makeatletter
+\newenvironment{assumption}[2]{
+	\renewcommand{\assumptionnumber}{Assumption #1#2}
+	\begin{assumption*}
+		\protected@edef\@currentlabel{#1#2}}
+	{\end{assumption*}}
+\makeatother
 
@@
 
-%%% End Ryan's definitions
+%%% End Ryan's template
 
@@
 
-\endlocaldefs
-
 \begin{document}
 
-\begin{frontmatter}
+\textcolor{red}{(TODO): Come back and edit this. For now it is garbled copy-paste gibberish.}
 
-\title{Supplementary Material For: Minimax-optimal Regression over Sobolev Spaces via Laplacian Eigenmaps with Neighborhood Graphs}
-%\thankstext{T1}{A sample of additional note to the title.}
+\section{Out-of-sample error}
+\label{sec:out_of_sample}
+We consider an estimator $\wh{f}$ defined only at the design points $\{X_1,\ldots,X_n\}$.
+Our goal is to define an extension to the entire domain $\mc{X}$, such that the resulting extension has small $L^2(P)$ norm.
+We propose a simple method, kernel smoothing, to do the job.
+The method can applied to any estimator defined at the design points, and we show that a smoothed version of our original estimator $\wh{f}$ has optimal $L^2(P)$ error.
+For simplicity, in this section we will stick to the flat Euclidean setting, where $(X_1,Y_1),\ldots,(X_n,Y_n)$ are observed according to Model~\ref{def:model_flat_euclidean}.
 
-\begin{aug}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Only one address is permitted per author. %%
-%% Only division, organization and e-mail is %%
-%% included in the address.                  %%
-%% Additional information can be included in %%
-%% the Acknowledgments section if necessary. %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author[A]{\fnms{Alden} \snm{Green}\ead[label=e1]{}},
-\author[A]{\fnms{Sivaraman} \snm{Balakrishnan}\ead[label=e1]{}}
-\and
-\author[A]{\fnms{Ryan J.} \snm{Tibshirani}\ead[label=e1]{}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Addresses                                %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\address[A]{Carnegie Mellon University, \printead{e1}}
-\end{aug}
+\paragraph{Extension by kernel smoothing.}
+We now formally define our approach to extension by kernel smoothing.
+For a kernel function $\psi(\cdot): [0,\infty) \to (-\infty,+\infty)$, bandwidth $h > 0$, and a distribution $Q$, the \emph{Nadaraya-Watson kernel smoother} $T_{Q,h}$ is given by
+\begin{equation*}
+\bigl(T_{Q,h}f)(x) := 
+\begin{dcases*}
+\frac{1}{d_{Q,h}(x)} \int_{\Omega} f(z)\psi\biggl(\frac{\|z - x\|}{h}\biggr) \,dQ(z), & \textrm{if $d_{Q,h}(x) > 0$,} \\
+0, &\textrm{otherwise,}
+\end{dcases*}
+\end{equation*}
+where $d_{Q,h}(x) := \int_{\Omega} \psi\bigl(\|z - x\|/h\bigr) \,dQ(z)$.
+For convenience, we will write $T_{n,h}f(x) := T_{P_n,h}f(x)$, and $d_{n,h}(x) := n \cdot d_{P_n,h}(x)$.
+We extend the Laplacian eigenmaps estimator by passing the kernel smoother $T_{n,h}$ over it, that is we consider the estimator $T_{n,h}\wh{f}$, which is defined at every $x \in \mc{X}$ (indeed, at every $x \in \Rd$).
+Note that ``extension'' here is a slight abuse of nomenclature, since $T_{n,h}\wh{f}$ and $\wh{f}$ may not agree in-sample.
 
-\end{frontmatter}
-% Main text
+\paragraph{Out-of-sample error of kernel smoothed Laplacian eigenmaps.}
 
-\section{Graph-dependent error bounds}
-\label{sec:fixed_graph_error_bounds}
-In this section, we adopt the fixed design perspective; or equivalently, condition on $X_i = x_i$ for $i = 1,\ldots,n$.
-Let $G = \bigl([n],W\bigr)$ be a fixed graph on $\{1,\ldots,n\}$ with Laplacian matrix $L = \sum_{k = 1}^{n}\lambda_k v_k v_k^{\top}$; the eigenvectors have unit empirical norm, $\|v_k\|_n^2 = 1$.
-The randomness thus all comes from the responses
+In Lemma~\ref{lem:kernel_smoothing_insample}, we consider an arbitrary estimator $\wc{f} \in \Reals^n$.
+We show that the out-of-sample error $\|T_{n,h}\wc{f} - f_0\|_P^2$ can be upper bounded by three terms--- (a constant times) the in-sample error $\|\wc{f} - f_0\|_n^2$, and variance and bias terms that arise naturally in the analysis of kernel smoothing over noiseless data.
+We shall assume the following conditions on $\psi$ and $h$.
+\begin{enumerate}[label=(K\arabic*)]
+	\setcounter{enumi}{2}
+	\item
+	\label{asmp:kernel}
+	The kernel function $\psi$ is supported on a subset of $[0,1]$. Additionally, $\psi$ is Lipschitz continuous on $[0,1]$, and is normalized so that
+	\begin{equation*}
+	\int_{-\infty}^{\infty} \psi(|z|) \,dz = 1.
+	\end{equation*}
+\end{enumerate}
+\begin{enumerate}[label=(P\arabic*)]
+	\setcounter{enumi}{6}
+	\item
+	\label{asmp:bandwidth}
+	For constants $c_0$ and $C_0$, the bandwidth parameter $h$ satisfies
+	\begin{equation*}
+	C_0\biggl(\frac{\log(1/h)}{n}\biggr)^{1/d} \leq h \leq c_0.
+	\end{equation*}
+\end{enumerate}
+\begin{lemma}
+	\label{lem:kernel_smoothing_insample}
+	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $f_0 \in H^1(\mc{X};M)$ and $p \in C^1(\mc{X})$.
+If the kernel smoothing estimator $T_{n,h}\wc{f}$ is computed with a kernel $\psi$ satisfying~\ref{asmp:kernel} and bandwidth $h$ satisfying~\ref{asmp:bandwidth}, it holds that
 	\begin{equation}
-\label{eqn:fixed_graph_regression_model}
-Y_i = f_{0}(x_i) + w_i
+	\label{eqn:kernel_smoothing_insample}
+	\|T_{n,h}\wc{f} - f_0\|_P^2 \leq C\biggl(\|\wc{f} - f_0\|_n^2 + \frac{1}{\delta} \cdot \frac{h^2}{nh^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{P,h}f_0 - f_0\|_P^2\biggr),
 	\end{equation}
-where the noise variables $w_i$ are independent $N(0,1)$.
-In the rest of this section, we will mildly abuse notation and write $f_0 = (f_0(x_1),\ldots,f_0(x_n)) \in \Reals^n$.
-We will also write ${\bf Y} = (Y_1,\ldots,Y_n)$.
+	with probability at least $1 - \delta - Ch^d\exp\{-Cnh^d\}$. 
+\end{lemma}
+Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data, by a factor of $h^2$.
+On the other hand the bias term is typical, and when $\psi$ is an order-$s$ kernel a standard analysis (see Lemma~\ref{lem:kernel_smoothing_bias}) shows that $\|T_{P,h}f_0 - f_0\|_P^2 \lesssim |f|_{H^s(\mc{X})}^2 h^{2s}$.
+\begin{enumerate}[label=(K\arabic*)]
+	\setcounter{enumi}{3}
+	\item
+	\label{asmp:ho_kernel}
+	The kernel function $\psi$ is an order-$s$ kernel, meaning it satisfies $\int_{-\infty}^{\infty} \psi(|z|) \,dz = 1$, and if $s \geq 2$,
+	\begin{equation*}
+	\int_{-\infty}^{\infty} z^j \psi(|z|) \,dz = 0 ~~\textrm{for}~j = 1,\ldots, s + d - 2, \quad \textrm{and}~ \int_{-\infty}^{\infty} z^{s + d - 1} \psi(|z|) \,dz < \infty. 
+	\end{equation*}
+\end{enumerate}
+Kernels satisfying~\ref{asmp:ho_kernel} can be constructed using linear combinations of (truncated) polynomials in a standard way; for instance, the boxcar kernel $\psi(z) = (1/2)\cdot\1\{|z| \leq 1\}$ is an order-$2$ kernel, and $\psi(z) = (9/26 - 15z^2/52)\cdot\1\{|z| \leq 1\}$ is an order-$3$ kernel.
 
-\subsection{Upper bound on Estimation Error of Laplacian eigenmaps}
+Choosing $h \asymp n^{-1/(2(s - 1) + d)}$ balances the kernel smoothing bias and variance terms in~\eqref{eqn:kernel_smoothing_insample}, and implies that for $f_0 \in H^s(\mc{X};M)$,
+\begin{equation}
+\label{eqn:kernel_smoothing_insample2}
+\|T_{n,h}\wc{f} - f_0\|_P^2 \leq C\biggl(\|\wc{f} - f_0\|_n^2 + \frac{M^2}{\delta}n^{-2s/(2(s - 1) + d)}\biggr).
+\end{equation}
+This analysis tells us that the additional error incurred by passing a kernel smoother over an in-sample estimator $\wc{f}$ is negligible compared to the minimax rate of estimation.
+Consequently, if $\wc{f}$ converges at the minimax rate in squared $L^2(P_n)$-norm, then $T_{n,h}\wc{f}$ will converge at the minimax rate in squared $L^2(P)$-norm.
+It follows immediately from Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} (when $s = 1$) or Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} (when $s > 1$) that $T_{n,h}\wh{f}$ achieves the optimal rate of convergence in $L^2(P)$.
 
-\begin{lemma}
-	\label{lem:fixed_graph_estimation}
-	For any integer $s > 0$, and any integer $0 \leq K \leq n$, the Laplacian eigenmaps estimator $\wh{f}$ satisfies
-	\begin{equation}
-	\label{eqn:fixed_graph_estimation}
-	\|\wh{f} - f_0\|_n^2 \leq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} + \frac{5K}{n};
-	\end{equation}
-	this is guaranteed if $K = 0$, and otherwise holds with probability at least $1 - \exp(-K)$ if $1 \leq K \leq n$. 
-\end{lemma}
-\paragraph{Proof (of Lemma~\ref{lem:fixed_graph_estimation}).}
-By the triangle inequality,
-\begin{equation}
-\label{pf:fixed_graph_estimation_1}
-\|\wh{f} - f_0\|_n^2 \leq 2\Bigl(\|\mathbb{E}\wh{f} - f_0\|_n^2 + \|\wh{f} - \mathbb{E}\wh{f}\|_n^2\Bigr).
-\end{equation}
-The first term in~\eqref{pf:fixed_graph_estimation_1} (approximation error) is non-random, since the design is fixed.
-The expectation $\mathbb{E}\wh{f} = \sum_{k = 1}^{K} \dotp{v_k}{f_0}_n v_k$, so that
+\begin{theorem}
+	\label{thm:laplacian_eigenmaps_estimation_out_of_sample}
+	Suppose Model~\ref{def:model_flat_euclidean}.
+There exist constants $c$, $C$, and $N$ that do not depend on $f_0$ or $n$ such that each the following statements hold with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\} - Ch^d\exp\{-cnh^d\}$,  for all $n \geq N$ and for any $\delta \in (0,1)$.
+	\begin{itemize}
+		\item If $f_0 \in H^1(\mc{X};M)$, the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_fo}, and the out-of-sample extension $T_{n,h}\wh{f}$ is computed with bandwidth $h = n^{-1/d}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel}, then
 		\begin{equation*}
-\|\mathbb{E}\wh{f} - f_0\|_n^2 = \Bigl\|\sum_{k = K + 1}^{n} \dotp{v_k}{f_0}_n v_k\Bigr\|_n^2 = \sum_{k = K + 1}^n \dotp{v_k}{f_0}_n^2.
+		\|T_{n,h}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2/(2 + d)}.
 		\end{equation*}
-In the above, the last equality relies on the fact that $v_k$ are orthonormal in $L^2(P_n)$. Using the fact that the eigenvalues are in increasing order, we obtain
+		\item If $f_0 \in H_0^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$ for some $s \in \mathbb{N}, s \geq 2$, and the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_ho}, and the out-of-sample extension $T_{n,h}\wh{f}$ is computed with bandwidth $h = n^{-1/(2(s - 1) + d)}$ and kernel $\psi$ that satisfies~\ref{asmp:kernel} and~\ref{asmp:ho_kernel}, then
 		\begin{equation*}
-\sum_{k = K + 1}^n \dotp{v_k}{f_0}_n^2 \leq \frac{1}{\lambda_{K + 1}^s} \sum_{k = K + 1}^n \lambda_k^s \dotp{v_k}{f_0}_n^2 \leq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s}.
+		\|T_{n,h}\wh{f} - f_0\|_P^2 \leq \frac{C}{\delta}M^2(M^2n)^{-2s/(2s + d)}.
 		\end{equation*}
+	\end{itemize}
+\end{theorem}
+Some remarks:
+\begin{itemize}
+	\item \emph{Data-driven parameter tuning}.
+Since $T_{n,h}\wh{f}$ is defined out-of-sample, we can use sample splitting or cross-validation methods to tune hyperparameters, which we could not do for the original estimator $\wh{f}$.
+For instance, we can (i) split the sample into two halves, (ii) use the first half to compute $T_{n,h}\wh{f}$ for various values of $\varepsilon$, $h$, and $K$, (iii) choose the optimal values of these three hyperparameters by minimizing error on the held out set.
+Practically speaking, cross-validation is one of the most common approaches to choosing hyperparameters.
+Theoretically, it is known that choosing hyper-parameters through sample splitting can result in estimators that optimally adapt to the order of regularity $s$ \citep{gyorfi2006}.
+In other words, it leads to estimators that are rate-optimal (up to $\log n$ factors), even when $s$ is unknown.
+We believe similar arguments should imply that $T_{n,h}\wh{f}$ is adaptive in this sense when $\varepsilon,h$ and $K$ are chosen by sample splitting, but the details should be worked out carefully.
 
-If $K = 0$, $\wh{f} = \Ebb{\wh{f}} = 0$, and the second term in~\eqref{pf:fixed_graph_estimation_1} is $0$.
-Otherwise the second   in~\eqref{pf:fixed_graph_estimation_1} (estimation error) is random.
-Observe that $\dotp{v_k}{\varepsilon}_n \overset{d}{=} Z_k/\sqrt{n}$, where $(Z_1,\ldots,Z_n) \sim N(0,I_{n \times n})$.
-Again using the orthonormality of the eigenvectors $v_k$, we have
-\begin{equation*}
-\|\wh{f} - \mathbb{E}\wh{f}\|_n^2 = \sum_{k = 1}^{K} \dotp{v_k}{\varepsilon}_n^2 \overset{d}{=} \frac{1}{n}\sum_{k = 1}^{K} Z_k^2.
-\end{equation*}
-Thus $\|\wh{f} - \mathbb{E}\wh{f}\|_n^2$ is equal to $1/n$ times a $\chi^2$ distribution with $K$ degrees of freedom. Consequently, it follows from a result of \citep{laurent00} that
-\begin{equation*}
-\Pbb\biggl(\|\wh{f} - \mathbb{E}\wh{f}\|_n^2 \geq \frac{K}{n} + 2\frac{\sqrt{K}}{n}\sqrt{t} + \frac{2t}{n}\biggr) \leq \exp(-t).
-\end{equation*}
-Setting $t = K$ completes the proof of the lemma.
+	% (AG 8/31/21): For now, I punted on proving formal results regarding sample-splitting. I can return later during a revision. 
+	\item \emph{Alternatives to extension by kernel smoothing}.
+There exist other approaches to extending a function $f$ from its evaluations $\{f(X_1),\ldots,f(X_n)\}$: for instance, minimum-norm interpolation in some RKHS or Banach space or 1-nearest neighbors regression.
+We consider extension by kernel smoothing because it is a simple and statistically optimal procedure that does not require any knowledge of the domain $\mc{X}$ or distribution $P$---as we have argued, this latter property is one of the main selling points of Laplacian eigenmaps as a tool for nonparametric regression.
+One potential drawback to kernel smoothing is that it can change the value of the estimate at data, meaning $\wh{f} \neq (\wc{f}(X_1),\ldots, \wc{f}(X_n))$.
+More sophisticated procedures based on Nystr\"{o}m extension of the eigenvectors $\{v_1,\ldots,v_K\}$~\citep{fowlkes2004} inherit the generality of kernel smoothing while resulting in genuine extrapolation; however, the statistical properties of these procedures are substantially more difficult to analyze.
+\end{itemize}
 
-\subsection{Upper bound on Testing Error of Laplacian Eigenmaps}
-
-Let $\wh{T} = \sum_{k = 1}^{K} \dotp{{\bf Y}}{v_k}_n^2$, and let $\varphi = \1\{\wh{T} \geq t_a\}$. In the following Lemma, we upper bound the Type I and Type II error of the test $\varphi$.
-
-\begin{lemma}
-	\label{lem:fixed_graph_testing}
-	Suppose we observe $(Y_1,x_1),\ldots,(Y_n,x_n)$ according to~\eqref{eqn:fixed_graph_regression_model}.
-	\begin{itemize}
-		\item If $f_0 = 0$, then $\Ebb_0[\varphi] \leq a$.
-		\item Suppose $f_0 \neq 0$ satisfies
-		\begin{equation}
-		\label{eqn:fixed_graph_testing_critical_radius}
-		\|f_0\|_n^2 \geq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} + \frac{\sqrt{2K}}{n}\biggl[2\sqrt{\frac{1}{a}} + \sqrt{\frac{2}{b}} + \frac{32}{bn}\biggr],
-		\end{equation}
-		for some $s \in \mathbb{N}\setminus \{0\}$. Then $\Ebb_{f_0}[1 - \phi] \leq b$.
-	\end{itemize}
-\end{lemma}
-\paragraph{Proof (of Lemma~\ref{lem:fixed_graph_testing}).}
-We first compute the expectation and variance of $\wh{T}$, then apply Chebyshev's inequality to upper bound the Type I and Type II error.
-
-\underline{\emph{Expectation}.}
-Recall that $\wh{T} = \sum_{k = 1}^{K} \dotp{Y}{v_k}_n^2$. Expanding the square gives
-\begin{equation*}
-\Ebb[\wh{T}] = \sum_{k = 1}^{K} \Ebb[\dotp{Y}{v_k}_n^2] = \sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 + \Ebb[2\dotp{f_0}{v_k}_n\dotp{\varepsilon}{v_k}_n + \dotp{\varepsilon}{v_k}_n^2] = \frac{K}{n} + \sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2.
-\end{equation*}
-Thus $\Ebb[\wh{T}] - t_a = \sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 - \sqrt{2K}/n \cdot \sqrt{1/a}$. Furthermore, it is a consequence of~\eqref{eqn:fixed_graph_testing_critical_radius} that 
-\begin{equation}
-\label{pf:fixed_graph_testing_1}
-\sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 - \frac{\sqrt{2K}}{n}\sqrt{1/a} \geq \|f_0\|_n^2 - \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} - \frac{\sqrt{2K}}{n}\sqrt{1/a} \geq \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \sqrt{\frac{2}{b}} + \frac{32}{bn}\biggr].
-\end{equation}
-
-\underline{\emph{Variance}.}
-Recall from the proof of Lemma~\ref{lem:fixed_graph_estimation} that $\dotp{\varepsilon}{v_k}_n \overset{d}{=} Z_k/\sqrt{n}$ for $(Z_1,\ldots,Z_n) \sim N(0,I_{n \times n})$.
-Expanding the square, and recalling that $\Cov[Z,Z^2] = 0$ for Gaussian random variables, we have that
-\begin{equation*}
-\Var\bigl[\dotp{{\bf Y}}{v_k}_n^2\bigr] = \Var\biggl[\frac{2}{n}\dotp{f_0}{v_k}_nZ_k + \frac{2}{n^2}Z_k^2\biggr] = \frac{4\dotp{f_0}{v_k}_n^2}{n} + \frac{2}{n^2}.
-\end{equation*}
-Moreover, since $\Cov[Z_k^2,Z_{\ell}^2] = 0$ for each $k = 1,\ldots,K$, we see that
-\begin{equation*}
-\Var\bigl[\wh{T}\bigr] = \sum_{k = 1}^{K} \Var\bigl[\dotp{{\bf Y}}{v_k}_n^2\bigr] = \frac{2K}{n^2} + \sum_{k = 1}^{K}\frac{4\dotp{f_0}{v_k}_n^2}{n}.
-\end{equation*}
-
-\underline{\emph{Bounds on Type I and Type II error}.}
-The upper bound on Type I error follows immediately from Chebyshev's inequality.
-
-The upper bound on Type II error also follows from Chebyshev's inequality.
-We observe that~\eqref{eqn:fixed_graph_testing_critical_radius} implies $\Ebb_{f_0}[\wh{T}] = t_a$, and apply Chebyshev's inequality to deduce
-\begin{equation*}
-\Pbb_{f_0}\bigl(\wh{T} < t_a\bigr) \leq \Pbb_{f_0}\Bigl(|\wh{T} - \Ebb_{f_0}[\wh{T}]|^2 > |\Ebb_{f_0}[\wh{T}] - t_a|^2\Bigr) \leq \frac{\Var\bigl[\wh{T}\bigr]}{\bigl[\Ebb_{f_0}[\wh{T}] - t_a\bigr]^2} = \frac{2K/n^2 + 4/n\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2}{\bigl[\Ebb_{f_0}[\wh{T}] - t_a\bigr]^2}.
-\end{equation*}
-Thus we have upper bounded the Type II error by the sum of two terms, each of which are no more than $1/(2b)$, as we now show.
-For the first term, after noting that~\eqref{pf:fixed_graph_testing_1} implies $\Ebb_{f_0}[\wh{T}] - t_a \geq \sqrt{2K}/n \cdot \sqrt{2/b}$, the upper bound follows:
-\begin{equation*}
-\frac{2K/n^2}{\bigl[\Ebb_{f_0}[\wh{T}] - t_a\bigr]^2} \leq \frac{b}{2}.
-\end{equation*}
-On the other hand, for the second term we use~\eqref{pf:fixed_graph_testing_1} in two ways: first to conclude that $\Ebb_{f_0}[\wh{T}] - t_a \geq 1/2 \cdot \sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2$, and second to obtain
-\begin{equation*}
-\frac{4\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2}{n\bigl[\Ebb_{f_0}[\wh{T}] - t_a\bigr]^2} \leq \frac{4\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2}{n\bigl(\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2/2\bigr)^2} \leq \frac{16}{n\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2} \leq \frac{b}{2}.
-\end{equation*}
-
-\section{Graph Sobolev semi-norm, flat Euclidean domain}
-\label{sec:graph_quadratic_form_euclidean}
-% AG: Hard-coded reference.
-In this section we prove Proposition~4. The proposition will follow from several intermediate results.
-\begin{enumerate}
-	\item~In Section~\ref{subsec:decomposition_graph_seminorm}, we show that
-	\begin{equation}
-	\label{pf:graph_seminorm_ho_1}
-	\dotp{L_{n,\varepsilon}^sf}{f}_n \leq \frac{1}{\delta} \dotp{L_{P,\varepsilon}^sf}{f}_{P} + \frac{C\varepsilon^2}{\delta n\varepsilon^{2 + d}}M^2.
-	\end{equation}
-	with probability at least $1 - 2\delta$.
-
-	We term the first term on the right hand side the \emph{non-local Sobolev semi-norm}, as it is a kernelized approximation to the Sobolev semi-norm $\dotp{\Delta_P^sf}{f}_{P}$.
-The second term on the right hand side is a pure bias term, which as we will see is negligible compared to the non-local Sobolev semi-norm as long as $\varepsilon \ll n^{-1/(2(s -1 + d))}$.
-	\item~In Section~\ref{subsec:approximation_error_nonlocal_laplacian}, we show that when $x$ is sufficiently in the interior of $\mc{X}$, then $L_{P,\varepsilon}^kf(x)$ is a good approximation to $\Delta_P^kf(x)$, as long as $f \in H^{s}(\mc{X})$ and $p \in C^{s - 1}(\mc{X})$ for some $s \geq 2k + 1$.
-	\item~In Section~\ref{subsec:boundary_behavior_nonlocal_laplacian}, we show that when $x$ is sufficiently near the boundary of $\mc{X}$, then $L_{P,\varepsilon}^kf(x)$ is close to $0$, as long as $f \in H_0^{s}(\mc{X})$ for some $s > 2k$.
-	\item~In Section~\ref{subsec:estimate_nonlocal_seminorm}, we use the results of the preceding two sections to show that if $f \in H_0^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$, there exists a constant $C$ which does not depend on $f$ such that
-	\begin{equation}
-	\label{pf:graph_seminorm_ho_2}
-	\dotp{L_{P,\varepsilon}^sf}{f}_{P} \leq CM^2.
-	\end{equation}
-\end{enumerate}
-Finally, in Section~\ref{subsec:integrals} we provide some assorted estimates used in Sections~\ref{subsec:decomposition_graph_seminorm}.
-
-\paragraph{Proof (of Proposition~4).}
-Proposition~4 follows immediately from~\eqref{pf:graph_seminorm_ho_1} and~\eqref{pf:graph_seminorm_ho_2}. \qed
-
-One note regarding notation: suppose a function $g \in H^{\ell}(U)$, where $\ell \in \mathbb{N}$ and $U$ is an open set.
-Let $V$ be another open set, compactly contained within $U$.
-Then we will use the notation $g \in H^{\ell}(V)$ to mean that the restriction $\restr{g}{V}$ of $g$ to $V$ belongs to $H^{\ell}(V)$.
-
-\subsection{Decomposition of graph Sobolev semi-norm}
-\label{subsec:decomposition_graph_seminorm}
-
-In Lemma~\ref{lem:graph_seminorm_bias}, we decompose the graph Sobolev semi-norm (a V-statistic) into an unbiased estimate of the non-local Sobolev semi-norm (a U-statistic), and a pure bias term.
-We establish that the pure bias term will be small (in expectation) relative to the U-statistic whenever $\varepsilon$ is sufficiently small.
-\begin{lemma}
-	\label{lem:graph_seminorm_bias}
-	For any $f \in L^2(\mc{X})$, the graph Sobolev semi-norm satisfies
-	\begin{equation}
-	\label{eqn:graph_seminorm_bias_1}
-	\dotp{L_{n,\varepsilon}^sf}{f}_{n} = U_{n,\varepsilon}^{(s)}(f) + B_{n,\varepsilon}^{(s)}(f),
-	\end{equation}
-	such that $\mathbb{E}[U_{n,\varepsilon}^{(s)}(f)] = (n - s - 1)!/n!
-\cdot \dotp{L_{P,\varepsilon}^sf}{f}_P$.
-If additionally $f \in H^1(\mc{X};M)$ and $\varepsilon \geq n^{-1/d}$, then the bias term $B_{n,\varepsilon}^{(s)}(f)$ satisfies
-	\begin{equation}
-	\label{eqn:graph_seminorm_bias_2}
-	\mathbb{E}\bigl[|B_{n,\varepsilon}^{(s)}(f)|\bigr] \leq \frac{C\varepsilon^2}{\delta n\varepsilon^{2 + d}}M^2.
-	\end{equation}
-\end{lemma}
-Then~\ref{pf:graph_seminorm_ho_1} follows immediately from Lemma~\ref{lem:graph_seminorm_bias}, by Markov's inequality.
-\paragraph{Proof (of Lemma~\ref{lem:graph_seminorm_bias}).}
-We begin by introducing some notation.
-We will use bold notation $\bj = (j_1,\ldots,j_s)$ for a vector of indices where $j_i \in [n]$ for each $i$.
-We write $[n]^s$ for the collection of all such vectors, and $(n)^s$ for the subset of such vectors with no repeated indices.
-Finally, we write $D_if$ for a kernelized difference operator,
-\begin{equation*}
-D_if(x) := \bigl(f(x) - f(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr),
-\end{equation*}
-and we let $D_{\bj}f(x) := \bigl(D_{j_1}\circ \cdots \circ D_{j_s}f\bigr)(x)$.
-
-With this notation in hand, it is easy to represent $\dotp{L_{n,\varepsilon}^sf}{f}_{n}$ as the sum of a U-statistic and a bias term,
-\begin{align*}
-\dotp{L_{n,\varepsilon}^sf}{f}_{n} & = \frac{1}{n} \sum_{i = 1}^{n} L_{n,\varepsilon}^sf(X_i) \cdot f(X_i) \\
-& = \underbrace{\frac{1}{n^{s + 1}\varepsilon^{s(d + 2)}} \sum_{{i\bf j} \in (n)^{s + 1}}D_{\bj}f(X_i) \cdot f(X_i)}_{=:U_{n,\varepsilon}^{(s)}(f)} + \underbrace{\frac{1}{n^{s + 1}\varepsilon^{s(d + 2)}} \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot f(X_i)}_{=:B_{n,\varepsilon}^{(s)}(f)}
-\end{align*}
-When the indices of $i\bj$ are all distinct, it follows straightforwardly from the law of iterated expectation that
-\begin{equation*}
-\mathbb{E}[ D_{\bj}f(X_i) \cdot f(X_i)] = \varepsilon^{s(d + 2)}\mathbb{E}[L_{P,\varepsilon}^sf(X_i) \cdot f(X_i)] = \dotp{L_{P,\varepsilon}^sf}{f}_{P}, 
-\end{equation*}
-which in turn implies $\mathbb{E}[U_{n,\varepsilon}^{(s)}(f)] = (n - s - 1)!/n! \cdot \dotp{L_{P,\varepsilon}^sf}{f}_P$.
-
-It remains to show~\eqref{eqn:graph_seminorm_bias_2}. By adding and subtracting $f(X_{\bj_1})$, we obtain by symmetry that
-\begin{equation*}
-\sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot f(X_i) = \frac{1}{2} \cdot \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot \bigl(f(X_i) - f(X_{\bj_1})\bigr),
-\end{equation*}
-and consequently
-\begin{equation*}
-\Ebb\Bigl[\sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot f(X_i)\Bigr] \leq \frac{1}{2} \cdot \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} \Ebb\Bigl[\bigl|D_{\bj}f(X_i)\bigr| \cdot \bigl|f(X_i) - f(X_{\bj_1})\bigr|\Bigr].
-\end{equation*}
-In Lemma~\ref{lem:graph_seminorm_bias2}, we show that if $f \in H^1(\mc{X};M)$, then for any $i\bj \in [n]^{s + 1}$ which contains a total of $k + 1$ distinct indices, 
-\begin{equation*}
-\Ebb\Bigl[\bigl|D_{\bj}f(X_i)\bigr| \cdot \bigl|f(X_i) - f(X_{\bj_1})\bigr|\Bigr] \leq C_1 \varepsilon^{2 + kd} M^2.
-\end{equation*}
-This shows us that the expectation of $|B_{n,\varepsilon}^s(f)|$ can bounded from above by the sum over several different terms, as follows:
-\begin{align*}
-\Ebb\Bigl[|B_{n,\varepsilon}^s(f)|\Bigr] & \leq C_1\frac{\varepsilon^2}{n\varepsilon^{2s}}M^2 \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} \frac{1}{(n\varepsilon^d)^s}  \varepsilon^{(|i\bj| - 1)d} \\
-& \leq C_1\frac{\varepsilon^2}{n\varepsilon^{2s}}M^2  \sum_{k = 1}^{s - 1} \frac{(n\varepsilon^d)^k}{(n\varepsilon^d)^s}n.
-\end{align*}
-Finally, we note that by assumption $n\varepsilon^d \geq 1$, so that in the above sum the factor of $(n\varepsilon^d)^k$ is largest when $k = s- 1$. We conclude that
-\begin{equation*}
-\Ebb\Bigl[|B_{n,\varepsilon}^s(f)|\Bigr] \leq C_1 (s - 1) \frac{\varepsilon^2}{n\varepsilon^{2s + d}}M^2,
-\end{equation*}
-which is the desired result.
-
-\subsection{Approximation error of non-local Laplacian}
-\label{subsec:approximation_error_nonlocal_laplacian}
-
-In this section, we establish the convergence $L_{P,\varepsilon}^kf \to \sigma_{\eta}^k\Delta_P^kf$ as $\varepsilon \to 0$.
-More precisely, we give an upper bound on the squared difference between $L_{P,\varepsilon}^kf$ and  $\sigma_{\eta}^k\Delta_P^kf$ as a function of $\varepsilon$.
-The bound holds for all $x \in \mc{X}_{k\varepsilon}$, and $f \in H^{s}(\mc{X})$, as long as $s \geq 2k + 1$.
-\begin{lemma}
-	\label{lem:approximation_error_nonlocal_laplacian}
-	Assume Model~1.
-Let $s \in \mathbb{N} \setminus \{0,1\}$, suppose that $f \in H^s(\mc{X};M)$, and if $s > 1$ suppose that $p \in C^{s - 1}(\mc{X})$.
-Let $L_{P,\varepsilon}$ be define with respect to a kernel $\eta$ that satisfies~(K1).
-Then there exist constants $C_1$ and $C_2$ that do not depend on $f$, such that each of the following statements hold.
-	\begin{itemize}
-		\item If $s$ is odd and $k = (s - 1)/2$, then
-		\begin{equation}
-		\label{eqn:approximation_error_nonlocal_laplacian_1}
-		\|L_{P,\varepsilon}^kf - \Delta_P^kf\|_{L^2(\mc{X}_{k\varepsilon})} \leq C_1 M \varepsilon
-		\end{equation}
-		\item If $s$ is even and $k = (s - 2)/2$, then
-		\begin{equation}
-		\label{eqn:approximation_error_nonlocal_laplacian_2}
-		\|L_{P,\varepsilon}^kf - \Delta_P^kf\|_{L^2(\mc{X}_{k\varepsilon})} \leq C_2 M \varepsilon^2.
-		\end{equation}
-	\end{itemize}
-\end{lemma}
-We remark that when $k = 1$ and $f \in C^3(\mc{X})$ or $C^4(\mc{X})$, statements of this kind are well known, and indeed stronger results---with $L^{\infty}(\mc{X})$ norm replacing $L^2(\mc{X})$ norm---hold.
-When dealing with the iterated Laplacian, and functions $f$ which are regular only in the Sobolev sense, the proof is somewhat more lengthy, but in result is similar in spirit.
-
-\paragraph{Proof (of Lemma~\ref{lem:approximation_error_nonlocal_laplacian}).}
-Throughout this proof, we shall assume that $f$ and $p$ are smooth functions, meaning they belong to $C^{\infty}(\mc{X})$.
-This is without loss of generality, since $C^{\infty}(\mc{X})$ is dense in both $H^s(\mc{X})$ and $C^{s - 1}(\mc{X})$, and since both sides of the inequalities~\eqref{eqn:approximation_error_nonlocal_laplacian_1} and~\eqref{eqn:approximation_error_nonlocal_laplacian_2} are continuous with respect to $\|\cdot\|_{H^s(\mc{X})}$ and $\|\cdot\|_{C^{s - 1}(\mc{X})}$ norms.
-
-We will actually prove a more general set of statements than contained in Lemma~\ref{lem:approximation_error_nonlocal_laplacian}, more general in the sense that they give estimates for all $k$, rather than simply the particular choices of $k$ given above.
-In particular, we will prove that the following two statements hold for any $s \in \mathbb{N}$ and any $k \in \mathbb{N} \setminus \{0\}$.
-\begin{itemize}
-	\item If $k \geq s/2$, then for every $x \in \mc{X}_{k\varepsilon}$, 
-	\begin{equation}
-	\label{pf:approximation_error_nonlocal_laplacian_0}
-	L_{P,\varepsilon}^kf(x) = g_s(x) \varepsilon^{s - 2k}
-	\end{equation}
-	for a function $g_s$ that satisfies
-	\begin{equation}
-	\label{pf:approximation_error_nonlocal_laplacian_0.5}
-	\|g_s\|_{L^2(\mc{X}_{k\varepsilon})} \leq C \|p\|_{C^{q}(\mc{X})}^k M 
-	\end{equation}
-	where $q = 1$ if $s =0$ or $s = 1$, and otherwise $q = s - 1$. 
-	\item If $k < s/2$, then for every $x \in \mc{X}_{k\varepsilon}$,
-	\begin{equation}
-	\label{pf:approximation_error_nonlocal_laplacian_1}
-	L_{P,\varepsilon}^kf(x) = \sigma_{\eta}^k \cdot \Delta_{P}^kf(x) + \sum_{j = 1}^{\floor{(s - 1)/2} - k} g_{2(j + k)}(x)\varepsilon^{2j} + g_{s}(x) \varepsilon^{s - 2k}.
-	\end{equation}
-	for functions $g_j$ that satisfy
-	\begin{equation}
-	\label{pf:approximation_error_nonlocal_laplacian_1.5}
-	\|g_j\|_{H^{s - j}(\mc{X}_{k\varepsilon})} \leq C \|p\|_{C^{s - 1}(\mc{X})}^k M.
-	\end{equation}
-\end{itemize}
-In the statement above, recall that $H^0(\mc{X}_{k\varepsilon}) = L^2(\mc{X}_{k\varepsilon})$.
-Additionally, note that we may speak of the pointwise behavior of derivatives of $f$ because we have assumed that $f$ is a smooth function.
-Observe that~\eqref{eqn:approximation_error_nonlocal_laplacian_1} follows upon taking $k = \floor{(s - 1)/2}$ in~\eqref{pf:approximation_error_nonlocal_laplacian_1}, whence we have
-\begin{equation*}
-\bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_{P}^kf(x)\bigr)^2 = \varepsilon^2 \bigl(g_s(x)\bigr)^2
-\end{equation*}
-for some $g_s \in L^2(\mc{X}_{k\varepsilon},C \cdot M \cdot \|p\|_{C^{s - 1}(\mc{X})})$, and integrating over $\mc{X}_{k\varepsilon}$ gives the desired result.
-\eqref{eqn:approximation_error_nonlocal_laplacian_2} follows from~\eqref{pf:approximation_error_nonlocal_laplacian_1} in an identical fashion.
-
-It thus remains establish~\eqref{pf:approximation_error_nonlocal_laplacian_1}, and~\eqref{pf:approximation_error_nonlocal_laplacian_0} which is an important part of proving~\eqref{pf:approximation_error_nonlocal_laplacian_1}.
-We will do so by induction on $k$.
-Note that throughout, we will let $g_j$ refer to functions which may change from line to line, but which always satisfy~\eqref{pf:approximation_error_nonlocal_laplacian_1.5}.
-
-\underline{\textit{Proof of~\eqref{pf:approximation_error_nonlocal_laplacian_0} and~\eqref{pf:approximation_error_nonlocal_laplacian_1}, base case.}}
-
-We begin with the base case, where $k = 1$.
-Again, we point out that although desired result is known when $s = 3$ or $s = 4$, and $f$ is regular in the H\"{o}lder sense, we require estimates for all $s \in \mathbb{N}$ when $f$ is regular in the Sobolev sense.
-
-When $s = 0$, the inequality~\eqref{pf:approximation_error_nonlocal_laplacian_0} is implied by Lemma~\ref{lem:l2estimate_nonlocal_laplacian}.
-When $s \geq 1$, we proceed using Taylor expansion.
-For any $x \in \mc{X}_{\varepsilon}$, we have that $B(x,\varepsilon) \subseteq \mc{X}$.
-Thus for any $x' \in B(x,\varepsilon)$, we may take an order $s$ Taylor expansion of $f$ around $x' = x$, and an order $q$ Taylor expansion of $p$ around $x' = x$, where $q = 1$ if $s = 1$, and otherwise $q = s - 1$.
-(See Section~\ref{subsec:taylor_expansion} for a review of the notation we use for Taylor expansions, as well as some properties that we make use of shortly.) This allows us to express $L_{P,\varepsilon}f(x)$ as the sum of three terms,
-\begin{align*}
-L_{P,\varepsilon}f(x) & = \frac{1}{\varepsilon^{d + 2}}\sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1}\frac{1}{j_1!j_2!}  \int_{\mc{X}} \bigl(d_x^{j_1}f\bigr)(x' - x) \bigl(d_x^{j_2}p\bigr)(x' - x) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \quad + \\
-& \quad \frac{1}{\varepsilon^{d + 2}}\sum_{j = 1}^{s - 1} \frac{1}{j!} \int_{\mc{X}} \bigl(d_x^jf\bigr)(x' - x)  r_{x'}^{q}(x;p) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \quad  + \\
-& \quad \frac{1}{\varepsilon^{d + 2}} \int_{\mc{X}} r_{x'}^j(x;f) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x').
-\end{align*}
-Here we have adopted the convention that $\sum_{j = 1}^{0} = 0$.
-
-Changing variables to $z = (x' - x)/\varepsilon$, we can rewrite the above expression as 
-\begin{align*}
-L_{P,\varepsilon}f(x) & = \frac{1}{\varepsilon^{2}}\sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1}\frac{\varepsilon^{j_1 + j_2}}{j_1!j_2!}  \int d_x^{j_1}f(z) d_x^{j_2}p(z) \eta\bigl(\|z\|\bigr) \,dz \quad + \\
-& \quad \frac{1}{\varepsilon^{2}} \sum_{j = 1}^{s - 1} \frac{\varepsilon^j}{j!} \int d_x^jf(z)  r_{zh + x}^{q}(x;p) \eta\bigl(\|z\|\bigr) \,dz \quad  + \\
-& \quad \frac{1}{\varepsilon^{2}} \int r_{zh + x}^j(x;f) \eta\bigl(\|z\|\bigr) p(zh + x)\,dz \\
-& := G_1(x) + G_2(x) + G_3(x).
-\end{align*}
-We now separately consider each of $G_1(x),G_2(x)$ and $G_3(x)$. We will establish that if $s = 1$ or $s = 2$, then $G_1(x) = 0$, and otherwise if $s \geq 3$ that
-\begin{equation*}
-G_1(x) = \sigma_{\eta}\Delta_Pf(x) + \sum_{j = 1}^{\floor{(s - 1)/2} - 1}g_{2(j + 1)}(x)\varepsilon^{2j} + g_{s}(x)\varepsilon^{s - 2}.
-\end{equation*}
-On the other hand, we will establish that if $s = 1$ then $G_2(x) = 0$, and otherwise for $s \geq 2$
-\begin{equation}
-\label{pf:approximation_error_nonlocal_laplacian_2}
-\|G_2\|_{L^2(\mc{X}_{\varepsilon})} \leq C \varepsilon^{s - 2} M \|p\|_{C^{s - 1}(\mc{X})};
-\end{equation}
-this same estimate will hold for $G_3$ for all $s \geq 1$. Together these will imply~\eqref{pf:approximation_error_nonlocal_laplacian_0} and~\eqref{pf:approximation_error_nonlocal_laplacian_1}.
-
-\emph{Estimate on $G_1(x)$.}
-If $s = 1$, then $s - 1 = 0$, and so $G_1(x) = 0$. We may therefore suppose $s \geq 2$. Recall that
-\begin{equation}
-G_1(x) = \sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1} \frac{\varepsilon^{j_1 + j_2 - 2}}{j_1!j_2!}  \underbrace{\int_{B(0,1)} d_x^{j_1}f(z) d_x^{j_2}p(z) \eta(\|z\|) \,dz}_{:= g_{j_1,j_2}(x)} \label{pf:approximation_error_nonlocal_laplacian_3}
-\end{equation}
-The nature of $g_{j_1,j_2}(x)$ depends on the sum $j_1 + j_2$.
-Since $d_x^{j_1}f d_x^{j_2}$ is an order $j_1 + j_2$ (multivariate) monomial, we have (see Section~\ref{subsec:taylor_expansion}) that whenever $j_1 + j_2$ is odd,
-\begin{equation*}
-g_{j_1,j_2}(x) = \int_{\mc{X}} d_x^{j_1}f(z) d_x^{j_2}p(z) \eta(\|z\|) \,dz = 0.
-\end{equation*}
-In particular this is the case when $j_1 = 1$ and $j_2 = 0$.
-Thus when $s = 2$,  $G_1(x) = g_{1,0}(x) = 0$.
-On the other hand if $s \geq 3$, then the lowest order terms in~\eqref{pf:approximation_error_nonlocal_laplacian_3} are those where $j_1 + j_2 = 2$, so that either $j_1 = 1$ and $j_2 = 1$, or $j_1 = 2$ and $j_2 = 0$.
-We have that
-\begin{align*}
-g_{1,1}(x) + \frac{1}{2}g_{2,0}(x) & = \int_{\mc{X}} d_x^{1}f(z) d_x^{1}p(z) \eta(\|z\|) \,dz + \frac{p(x)}{2} \int_{\mc{X}} d_x^{2}f(z) \eta(\|z\|) \,dz \\
-& = \sum_{i_1 = 1}^{d} \sum_{i_2 = 1}^{d} D^{e_{i_1}}f(x) D^{e_{i_2}}p(x) \int_{\mc{X}} z^{e_{i_1} + e_{i_2}} \eta(\|z\|) \,dz + \frac{p(x)}{2} \sum_{i_1 = 1}^{d} \sum_{i_2 = 1}^{d}D^{e_{i_2}+e_{i_2}}f(x)\int_{\mc{X}} z^{e_{i_1} + e_{i_2}} \eta(\|z\|) \,dz\\
-& = \sum_{i = 1}^{d} D^{e_{i}}f(x) D^{e_{i}}p(x) \int_{\mc{X}} z^2 \eta(\|z\|) \,dz + \frac{p(x)}{2} \sum_{i = 1}^{d} D^{2e_{i}}f(x)\int_{\mc{X}} z^2 \eta(\|z\|) \,dz\\ 
-& = \sigma_{\eta}\Delta_Pf(x),
-\end{align*}
-which is the leading term order term.
-Now it remains only to deal with the higher-order terms, where $j_1 + j_2 > 2$, and where it suffices to show that each function $g_{j_1,j_2}$ satisfies~\eqref{pf:approximation_error_nonlocal_laplacian_1.5} for $j = \min\{j_1 + j_2 - 2,s - 2\}$.
-It is helpful to write $g_{j_1,j_2}$ using multi-index notation,
-\begin{align*}
-g_{j_1,j_2}(x) = \sum_{|\alpha_1| = j_1} \sum_{|\alpha_2| = j_2} D^{\alpha_1}f(x) D^{\alpha_2}p(x) \int_{B(0,1)} z^{\alpha_1 + \alpha_2} \eta(\|z\|) \,dz,
-\end{align*}
-where we note that $|\int_{B(0,1)} z^{\alpha_1 + \alpha_2} \eta(\|z\|) \,dz| < \infty$ for all $\alpha_1, \alpha_2$, by the assumption that $\eta$ is Lipschitz on its support.
-Finally, by H\"{o}lder's inequality we have that
-\begin{align*}
-\|D^{\alpha_1}f D^{\alpha_2}p\|_{H^{s - (j + 2)}(\mc{X})} & \leq \|D^{\alpha_1}f\|_{H^{s - (j + 2)}(\mc{X})} \|D^{\alpha_2}p\|_{C^{s - (j + 2)}(\mc{X})} \\
-& \leq \|D^{\alpha_1}f\|_{H^{s - j_1}(\mc{X})} \|D^{\alpha_2}p\|_{C^{s - (j_2 + 1)}(\mc{X})} \\
-& \leq M \cdot \|p\|_{C^{s - 1}(\mc{X})},
-\end{align*}
-and summing over all $|\alpha_1| = j_1$ and $|\alpha_2| = j_2$ establishes that $g_{j_1,j_2}$ satisfies~\eqref{pf:approximation_error_nonlocal_laplacian_1.5}.
-
-\emph{Estimate on $G_2(x)$.}
-Note immediately that $G_2(x) = 0$ if $s = 1$.
-Otherwise if $s \geq 2$, then $q = s - 1$.
-Recalling that $|r_{x + z\varepsilon}^{s - 1}(x; p)| \leq C\varepsilon^{s - 1}\|p\|_{C^{s - 1}(\mc{X})}$ for any $z \in B(0,1)$, and that $d_x^jf(\cdot)$ is a $j$-homogeneous function, we have that
-\begin{align}
-|G_2(x)| & \leq \sum_{j = 1}^{s - 1} \frac{\varepsilon^{j - 2}}{j!}\int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(z)\Bigr| \cdot |r_{x + z\varepsilon}^{s - 1}(x;p)| \cdot \eta(\|z\|) \,dz \nonumber \\
-& \leq C\varepsilon^{s - 2}\|p\|_{C^{s - 1}(\mc{X})} \sum_{j = 1}^{s - 1} \frac{1}{j!} \int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(z)\Bigr| \cdot \eta(\|z\|) \,dz \label{pf:approximation_error_nonlocal_laplacian_4}.
-\end{align}
-Furthermore, for each $j = 1,\ldots,s - 1$ convolution of $d_x^jf$ with $\eta$ only decreases the $L^2(\mc{X}_{\varepsilon})$ norm, meaning
-\begin{equation}
-\label{pf:approximation_error_nonlocal_laplacian_5}
-\begin{aligned}
-\int_{\mc{X}_{\varepsilon}} \biggl(\int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(z)\Bigr| \cdot \eta(\|z\|) \,dz\biggr)^2 \,dx & \leq \int_{\mc{X}_{\varepsilon}} \biggl(\int_{B(0,1)} \Bigl|\bigl(d_x^jf\bigr)(z)\Bigr|^2 \eta(\|z\|)\,dz \biggr) \cdot \biggl(\int_{B(0,1)} \eta(\|z\|) \,dz \biggr) \,dx \\
-& \leq \int_{B(0,1)} \int_{\mc{X}_{\varepsilon}} \Bigl[\bigl(d^jf\bigr)(x)\Bigr]^2 \eta(\|z\|) \,dx  \,dz \\
-& \leq \|d^jf\|_{L^2(\mc{X_{\varepsilon}})}^2.
-\end{aligned}
-\end{equation}
-In the above, we have used both that $|d_x^jf(z)| \leq |d^jf(x)|$ for all $z \in B(0,1)$, and that the kernel is normalized so that $\int \eta(\|z\|) \,dz = 1$. 
-Combining this with~\eqref{pf:approximation_error_nonlocal_laplacian_4}, we conclude that
-\begin{align*}
-\int_{\mc{X}_{\varepsilon}} |G_2(x)|^2 \,dx & \leq C \Bigl(\varepsilon^{s - 2}\|p\|_{C^{s - 1}(\mc{X})}\Bigr)^2 \sum_{j = 1}^{s - 1} \int_{\mc{X}_{\varepsilon}}\biggl(\frac{1}{j!} \int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(z)\Bigr| \cdot \Bigl|\eta(\|z\|)\Bigr| \,dz\biggr)^2 \,dx \\
-& \leq C \Bigl(\varepsilon^{s - 2}\|p\|_{C^{s - 1}(\mc{X})}\Bigr)^2 \sum_{j = 1}^{s - 1} \|d^ju\|_{L^2(\mc{X_{\varepsilon}})}^2,
-\end{align*}
-establishing the desired estimate.
-
-\emph{Estimate on $G_3(x)$.}
-Applying the Cauchy-Schwarz inequality, we deduce a pointwise upper bound on $|G_3(x)|^2$,
-\begin{align*}
-|G_3(x)|^2 & \leq \biggl(\frac{p_{\max}}{\varepsilon^2}\biggr)^2 \cdot \biggl(\int_{B(0,1)} \bigl|r_{x + \varepsilon z}^s(x;u)\bigr|^2 \eta(\|z\|)\,dz\biggr) \cdot \biggl(\int_{B(0,1)} \eta(\|z\|) \,dz\biggr) \\
-& \leq \biggl(\frac{p_{\max}}{\varepsilon^2}\biggr)^2 \int_{B(0,1)} \bigl|r_{x + \varepsilon z}^s(x;u)\bigr|^2 \eta(\|z\|) \,dz.
-\end{align*}
-Applying this pointwise over all $x \in \mc{X}_{\varepsilon}$ and integrating, we obtain
-\begin{align*}
-\int_{\mc{X}_{\varepsilon}} |G_3(x)|^2 \,dx & \leq \biggl(\frac{p_{\max}}{\varepsilon^2}\biggr)^2 \int_{\mc{X}_{\varepsilon}} \int_{B(0,1)} \bigl|r_{x + \varepsilon z}^s(x;f)\bigr|^2 \eta(\|z\|) \,dz \,dx \\
-& = \biggl(\frac{p_{\max}}{\varepsilon^2}\biggr)^2 \int_{B(0,1)} \int_{\mc{X}_{\varepsilon}} \bigl|r_{x + \varepsilon z}^s(x;f)\bigr|^2 \eta(\|z\|) \,dx \,dz \\
-& \leq \biggl(\frac{p_{\max}\varepsilon^s}{\varepsilon^2}\biggr)^2  \|d^sf\|_{L^2(\mc{X}_{\varepsilon})}^2,
-\end{align*}
-with the last inequality following from~\eqref{eqn:sobolev_remainder_term}.
-Noting that $p_{\max} = \|p\|_{C^0(\mc{X})} \leq \|p\|_{C^{s - 1}(\mc{X})}$, we see that this is a sufficient bound on $\|G_3\|_{L^2(\mc{X}_{\varepsilon})}$.
-
-\underline{\textit{Proof of~\eqref{pf:approximation_error_nonlocal_laplacian_0} and~\eqref{pf:approximation_error_nonlocal_laplacian_1}, induction step.}}
-We now assume that~\eqref{pf:approximation_error_nonlocal_laplacian_0} and~\eqref{pf:approximation_error_nonlocal_laplacian_1} hold for all order up to some $k$, and show that they then hold for order $k + 1$ as well.
-The proof is relatively straightforward, once we introduce a bit of notation.
-Namely, for any $\ell,j \in \mathbb{N}$ such that $1 \leq j \leq \ell \leq$, we will use $g_j^{\ell}$ to refer to a function satisfying
-\begin{equation}
-\label{pf:approximation_error_nonlocal_laplacian_6}
-\|g_j^{\ell}\|_{H^{\ell - j}(\mc{X}_{(k + 1)\varepsilon})} \leq C \|p\|_{C^{q}(\mc{X})}^{k + 1} M.
-\end{equation}
-Note that $g_j^{\ell}(x) = g_{(s - \ell) + j}(x)$, so that $g_j^{s}(x) = g_j(x)$.
-As before, the functions $g_j^{\ell}$ may change from line to line, but will always satisfy~\eqref{pf:approximation_error_nonlocal_laplacian_6}.
-We immediately illustrate the purpose of this notation.
-Suppose $g \in H^{\ell}(\mc{X}_{k\varepsilon}; C \|p\|_{C^{q}(\mc{X})}^k M)$ for some $\ell \leq s$.
-If $\ell \leq 2$, then by the inductive hypothesis, it follows that for any $x \in \mc{X}_{(k + 1)\varepsilon}$
-\begin{equation}
-\label{pf:approximation_error_nonlocal_laplacian_7}
-L_{P,\varepsilon}g(x) = g_{\ell}^{\ell}(x) \varepsilon^{\ell - 2}.
-\end{equation} 
-On the other hand if $2 < \ell \leq s$, then by the inductive hypothesis, it follows that for any $x \in \mc{X}_{(k + 1)\varepsilon}$,
-\begin{equation}
-\label{pf:approximation_error_nonlocal_laplacian_8}
-L_{P,\varepsilon}g(x) = \sigma_{\eta} \Delta_Pg(x) + \sum_{j = 1}^{\floor{(\ell - 1)/2} - 1} g_{2j + 2}^{\ell}(x) \varepsilon^{2j} + g_{\ell}^{\ell}(x) \varepsilon^{\ell - 2}.
-\end{equation}
-
-\emph{Proof of \eqref{pf:approximation_error_nonlocal_laplacian_0}.} If $s \leq 2(k + 1)$, then by the inductive hypothesis it follows that for all $x \in \mc{X}_{k\varepsilon}$, we have $L_{P,\varepsilon}^kf(x) = g_{s}(x) \cdot \varepsilon^{s - 2k}$, for some $g_s \in L^2(\mc{X}_{k\varepsilon}, C\|p\|_{C^{s - 1}(\mc{X})}^k M)$.
-Note that we may know more about $L_P^kf(x)$ than simply that it is bounded in $L^2$-norm, but a bound in $L^2$-norm suffices.
-In particular, from such a bound along with~\eqref{pf:approximation_error_nonlocal_laplacian_7} we deduce that for any $x \in \mc{X}_{(k + 1)\varepsilon}$,
-\begin{equation}
-\label{pf:approximation_error_nonlocal_laplacian_8.5}
-L_{P,\varepsilon}^{k + 1}f(x) = \bigl(L_{P,\varepsilon} \circ L_{P,\varepsilon}^k f)(x)= L_{P,\varepsilon} g_s(x)\varepsilon^{s - 2k} = g_{s}^{s}(x) \varepsilon^{s - 2(k + 1)},
-\end{equation}
-establishing~\eqref{pf:approximation_error_nonlocal_laplacian_0}.
-
-\emph{Proof of \eqref{pf:approximation_error_nonlocal_laplacian_1}.} If $s > 2(k + 1)$, then by the inductive hypothesis we have that for all $x \in \mc{X}_{k\varepsilon}$, 
-\begin{equation*}
-L_{P,\varepsilon}^kf(x) = \sigma_{\eta}^k \Delta_P^kf(x) + \sum_{j = 1}^{\floor{(s - 1)/2} - k} g_{2(j + k)}(x) \varepsilon^{2j} + g_s(x) \varepsilon^{s - 2k}.
-\end{equation*}
-Thus for any $x \in \mc{X}_{(k + 1)\varepsilon}$, 
-\begin{equation*}
-L_{P,\varepsilon}^{k + 1}f(x) = \bigl(L_{P,\varepsilon} \circ L_{P,\varepsilon}^k f\bigr)(x) = \sigma_{\eta}^k L_{P,\varepsilon}\Delta_P^kf(x) + \sum_{j = 1}^{\floor{(s - 1)/2} - k} L_{P,\varepsilon}g_{2(j + k)}(x) \varepsilon^{2j} + L_{P,\varepsilon}g_s(x) \varepsilon^{s - 2k}
-\end{equation*}
-There are three terms on the right hand side of this equality, and we now analyze each separately.
-\begin{enumerate}
-	\item Noting that $\Delta_P^kf \in H^{s - 2k}(\mc{X}; C\|p\|_{C^{s - 1}(\mc{X})}^kM)$, we use~\eqref{pf:approximation_error_nonlocal_laplacian_8} to derive that
-	\begin{align}
-	L_{P,\varepsilon}\Delta_P^kf(x) & = \sigma_{\eta} \Delta_P^{k + 1}f(x) + \sum_{j = 1}^{(s - 2k - 1)/2 - } g_{2j + 2}^{s - 2k}(x)\varepsilon^{2j} + g_{s - 2k}^{s - 2k}(x) \varepsilon^{s - 2k - 2} \nonumber \\
-	& = \sigma_{\eta} \Delta_P^{k + 1}f(x) + \sum_{j = 1}^{(s - 1)/2 - (k + 1)} g_{2(k + 1 + j)}(x)\varepsilon^{2j} + g_{s}(x) \varepsilon^{s - 2(k + 1)}, \label{pf:approximation_error_nonlocal_laplacian_9}
-	\end{align}
-	where in the second equality we have simply used the fact $g_j^{\ell}(x) = g_{(s - \ell) + j}(x)$ to rewrite the equation.
-	\item Suppose $j < \floor{(s - 1)/2} - k$. Then we use~\eqref{pf:approximation_error_nonlocal_laplacian_8} to derive that
-	\begin{align*}
-	L_{P,\varepsilon}g_{2(j + k)}(x) & = \sigma_{\eta}\Delta_P g_{2(j + k)}(x) + \sum_{i = 1}^{\floor{(s - 2j - 2k - 1)/2} - 1} g_{2(i + 1)}^{s - 2(j + k)}(x)\varepsilon^{2i} + g_{s - 2(j + k)}^{s - 2(j + k)}(x) \varepsilon^{s - 2(j + k + 1)} \\
-	& = g_{2(j + k + 1)}(x) + \sum_{i = 1}^{\floor{(s - 1)/2} - (j + k + 1)} g_{2(i + j + k + 1)}(x)\varepsilon^{2i} + g_{s}(x) \varepsilon^{s - 2(j + k + 1)},
-	\end{align*}
-	where in the second equality we have again used $g_j^{\ell}(x) = g_{(s - \ell) + j}(x)$, and also written $\sigma_{\eta} \Delta_Pf = g_{2}^{s - 2(j + k)} = g_{2(j + k + 1)}$, since the particular dependence on the Laplacian $\Delta_P$ will not matter.
-From here, multiplying by $\varepsilon^{2j}$, we conclude that
-	\begin{align}
-	\varepsilon^{2j} L_{P,\varepsilon}g_{2(j + k)}(x) & = g_{2(j + k + 1)}(x) \varepsilon^{2j} + \sum_{i = 1}^{\floor{(s - 1)/2} - (j + k + 1)} g_{2(i + j + k + 1)}(x)\varepsilon^{2(i + j)} + g_{s}(x) \varepsilon^{s - 2(k + 1)} \nonumber \\
-	& = g_{2(j + k + 1)}(x) \varepsilon^{2j} + \sum_{m = 1}^{\floor{(s - 1)/2} - (k + 1)}  g_{2(m + k + 1)}(x)\varepsilon^{2m} + g_{s}(x) \varepsilon^{s - 2(k + 1)} \label{pf:approximation_error_nonlocal_laplacian_10},
-	\end{align}
-	with the second equality following upon changing variables to $m = i + j$.
-
-	On the other hand if $j = \floor{(s - 1)/2} - k$, then the calculation is much simpler,
-	\begin{equation}
-	\label{pf:approximation_error_nonlocal_laplacian_11}
-	\varepsilon^{2j} L_{P,\varepsilon}g_{2(j + k)}(x) = g_{s - 2(j + k)}^{s - 2(j + k)}(x) \varepsilon^{2j} \varepsilon^{s - 2(j + k) - 2} = g_s(x) \varepsilon^{s - 2(k + 1)}.
-	\end{equation}
-	\item Finally, it follows immediately from~\eqref{pf:approximation_error_nonlocal_laplacian_8} that
-	\begin{equation}
-	\label{pf:approximation_error_nonlocal_laplacian_12}
-	L_{P,\varepsilon}g_s(x) \varepsilon^{s - 2k} = g_{s}(x) \varepsilon^{s - 2(k + 1)}.
-	\end{equation}
-\end{enumerate}
-Plugging~\eqref{pf:approximation_error_nonlocal_laplacian_9}-\eqref{pf:approximation_error_nonlocal_laplacian_12} back into~\eqref{pf:approximation_error_nonlocal_laplacian_8.5} proves the claim.
-
-\subsection{Boundary behavior of non-local Laplacian}
-\label{subsec:boundary_behavior_nonlocal_laplacian}
-
-In Lemma~\ref{lem:approximation_error_nonlocal_laplacian_boundary}, we establish that if $f$ is Sobolev smooth of order $s > 2k$ and zero-trace, then near the boundary of $\mc{X}$ the non-local Laplacian $L_{P,\varepsilon}^kf$ is close to $0$ in the $L^2$-sense.
-\begin{lemma}
-	\label{lem:approximation_error_nonlocal_laplacian_boundary}
-	Assume Model~1. Let $s,k \in \mathbb{N}$. Suppose that $f \in H_0^{s}(\mc{X};M)$. Then there exist numbers $c,C > 0$ that do not depend on $M$, such that for all $\varepsilon < c$, 
-	\begin{equation*}
-	\|L_{P,\varepsilon}^kf\|_{L^2(\partial_{k\varepsilon}\mc{X})}^2 \leq C \varepsilon^{2(s - 2k)}M^2.
-	\end{equation*}
-\end{lemma}
-
-\paragraph{Proof (of Lemma~\ref{lem:approximation_error_nonlocal_laplacian_boundary})}
-Applying Lemma~\ref{lem:l2estimate_nonlocal_laplacian}, we have that
-\begin{equation*}
-\|L_{P,\varepsilon}^kf\|_{L^2(\partial_{k\varepsilon}(\mc{X}))}^2 \leq \frac{(Cp_{\max})^{2}}{\varepsilon^4} \|L_{P,\varepsilon}^{k - 1}f\|_{L^2(\partial_{k\varepsilon}(\mc{X}))}^2 \leq \cdots \leq \frac{(Cp_{\max})^{2}}{\varepsilon^{4k}} \|f\|_{L^2(\partial_{k\varepsilon}(\mc{X}))}^2
-\end{equation*}
-Thus it remains to show that for all $\varepsilon < c$,
-\begin{equation}
-\label{pf:approximation_error_nonlocal_laplacian_boundary_0}
-\|f\|_{L^2(\partial_{k\varepsilon}(\mc{X}))}^2 = \int_{\partial_{k\varepsilon}(\mc{X})} \bigl(f(x)\bigr)^2 \,dx \leq C_1 \varepsilon^{2s} \|f\|_{H^s(\mc{X})}^2.
-\end{equation}
-We will build to~\eqref{pf:approximation_error_nonlocal_laplacian_boundary_0} by a series of intermediate steps, following the same rough structure as the proof of Theorem 18.1 in \citet{leoni2017}.
-For simplicity, we will take $k = 1$; the exact same proof applies to the general case upon assuming $\varepsilon < c/k$.
-
-\underline{\textit{Step 1: Local Patch.}}
-To begin, we assume that for some $c_0 > 0$ and a Lipschitz mapping $\phi: \Reals^{d - 1} \to [-c_0,c_0]$, we have that $f \in C_c^{\infty}(U_{\phi}(c_0))$, where 
-\begin{equation*}
-U_{\phi}(c_0) = \Bigl\{y \in Q(0,c_0): \phi(y_{-d}) \leq y_d\Bigr\}, 
-\end{equation*}
-and here $Q(0,c_0)$ is the $d$-dimensional cube of side length $c_0$, centered at $0$.
-We will show that for all $0 < \varepsilon < c_0$, and for the tubular neighborhood $V_{\phi}(\varepsilon) = \{y \in Q(0,c_0): \phi(y_{-d}) \leq y_d \leq \phi(y_{-d}) + \varepsilon\}$, we have that
-\begin{equation*}
-\int_{V_{\phi}(\varepsilon)} |f(x)|^2 \,dx \leq C\varepsilon^{2s} \|f\|_{H^s(U_{\phi}(c_0))}^2.
-\end{equation*}
-For a given $y = (y',y_d) \in V_{\phi}(\varepsilon)$, let $y_0 = (y',\phi(y'))$.
-Taking the Taylor expansion of $f(y)$ around $y = y_0$ because $u$ is compactly supported in $V_{\phi}$ it follows that,
-\begin{align*}
-f(y) & = f(y_0) + \sum_{j = 1}^{s - 1} \frac{1}{j!} D^{je_d}f(y_0) \bigl(y_d - \phi(y')\bigr)^j + \frac{1}{(s - 1)!}\int_{\phi(y')}^{y_d} (1 - t)^{s - 1} D^{se_d}f(y',z) \bigl(y_d - z\bigr)^{s - 1} \,dz \Longrightarrow\\
-|f(y)| & \leq C\varepsilon^{s - 1}\int_{\phi(y')}^{y_d} \bigl|D^{se_d}f(y',z)\bigr| \,dz. 
-\end{align*}
-Consequently, by squaring both sides and applying Cauchy-Schwarz, we have that
-\begin{equation*}
-|f(y)|^2 \leq C\varepsilon^{2(s - 1)} \biggl(\int_{\phi(y')}^{y_d} \bigl|D^{se_d}f(y',z)\bigr| \,dz\biggr)^2 \leq C\varepsilon^{2s - 1} \int_{\phi(y')}^{y_d} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz.
-\end{equation*}
-Applying this bound for each $y \in V_{\phi}(\varepsilon)$, and then integrating, we obtain
-\begin{align}
-\int_{V_{\phi}(\varepsilon)} |f(y)|^2 \,dy & \leq \int_{Q_{d - 1}(c_0)} \int_{\phi(y')}^{\phi(y') + \varepsilon} |f(y',y_d)|^2 \,dy_d \,dy' \nonumber \\
-& \leq C\varepsilon^{2s - 1}\int_{Q_{d - 1}(c_0)}  \int_{\phi(y')}^{\phi(y') + \varepsilon} \int_{\phi(y')}^{y_d} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz \,dy_d \,dy' \label{pf:approximation_error_nonlocal_laplacian_boundary_1}
-\end{align}
-where we have written $Q_{d - 1}(0,c_0)$ for the $d - 1$ dimensional cube of side length $c_0$, centered at $0$. Exchanging the order of the inner two integrals then gives
-\begin{align*}
-\int_{\phi(y')}^{\phi(y') + \varepsilon} \int_{\phi(y')}^{y_d} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz \,dy_d & = \int_{\phi(y')}^{\phi(y') + \varepsilon} \int_{z}^{\varepsilon} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dy_d \,dz \\
-& \leq C \varepsilon \int_{\phi(y')}^{\phi(y') + \varepsilon} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz \\
-& \leq C \varepsilon \int_{\phi(y')}^{c_0} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz.
-\end{align*}
-Finally, plugging back into~\eqref{pf:approximation_error_nonlocal_laplacian_boundary_1}, we conclude that
-\begin{equation*}
-\int_{V_{\phi}(\varepsilon)} |f(y)|^2 \,dy \leq C \varepsilon^{2s} \int_{Q_{d - 1}(0,c_0)} \int_{\phi(y')}^{c_0} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz \,dy' \leq C \varepsilon^{2s} |u|_{H^s(U_{\phi}(c_0))}^2.
-\end{equation*}
-
-\underline{\textit{Step 2: Rigid motion of local patch.}} Now, suppose that at a point $x_0 \in \partial \mc{X}$, there exists a rigid motion $T: \Rd \to \Rd$ for which $T(x_0) = 0$, and a number $C_0$ such that for all $\varepsilon \cdot C_0 \leq c_0$,
-\begin{equation*}
-T\bigl(Q_{T}(x_0,c_0) \cap \partial_{\varepsilon}\mc{X}\bigr) \subseteq V_{\phi}\bigl(C_0\varepsilon\bigr) \quad\textrm{and}\quad T\bigl(Q_T(x_0,c_0) \cap \mc{X}\bigr) = U_{\phi}(c_0).
-\end{equation*}
-Here $Q_{T}(x_0,c_0))$ is a (not necessarily coordinate-axis-aligned) cube of side length $c_0)$, centered at $x_0$.
-Define $v(y) := f(T^{-1}(y))$ for $y \in U_{\phi}(c_0)$.
-If $u \in C_c^{\infty}(\mc{X})$, then $v \in C_c^{\infty}(U_{\phi}(c_0))$, and moreover $\|v\|_{H^s(U_{\phi}(c_0))}^2 = \|f\|_{H^s(Q_{T}(x_0,c_0) \cap \mc{X})}^2$.
-Therefore, using the upper bound that we derived in Step 1,
-\begin{equation*}
-\int_{V_{\phi}(C_0 \cdot \varepsilon)} |v(y)|^2 \,dy \leq C \varepsilon^{2s} \|v\|_{H^s(U_{\phi}(c_0))}^2,
-\end{equation*}
-we conclude that
-\begin{align*}
-\int_{Q_{T}(x_0,c_0) \cap \partial_{\varepsilon}\mc{X}} |f(x)|^2 \,dx & = \int_{T(Q_T(x_0,c_0)) \cap \partial_{\varepsilon}\mc{X})} |v(y)|^2 \,dy \\
-& \leq \int_{V_{\phi}(C_0 \cdot \varepsilon)} |v(y)|^2 \,dy \\
-& \leq C \varepsilon^{2s} \|v\|_{H^s(U_{\phi}(c_0))}^2 = C \varepsilon^{2s} \|f\|_{H^s(Q_{T}(x_0,c_0)) \cap \mc{X})}^2 \leq C \varepsilon^{2s} \|f\|_{H^s(\mc{X})}^2.
-\end{align*}
-
-\underline{\textit{Step 3: Lipschitz domain}}.
-Finally, we deal with the case where $\mc{X}$ is assumed to be an open, bounded subset of $\Rd$, with Lipschitz boundary.
-In this case, at every $x_0 \in \partial \mc{X}$, there exists a rigid motion $T_{x_0}: \Rd \to \Rd$ such that $T_{x_0}(x_0) = 0$, a number $c_0(x_0)$, a Lipschitz function $\phi_{x_0}:\Reals^{d - 1} \to [-c_0,c_0]$, and a number $C_0(x_0)$, such that for all $\varepsilon \cdot C_0(x_0) \leq c_0(x_0)$,
-\begin{equation*}
-T\bigl(Q_{T}(x_0,c_0(x_0)) \cap \partial_{\varepsilon}\mc{X}\bigr) \subseteq V_{\phi}\bigl(C_0(x_0) \cdot \varepsilon\bigr) \quad\textrm{and}\quad T\bigl(Q_T(x_0,c_0(x_0)) \cap \mc{X}\bigr) = U_{\phi}(c_0(x_0)).
-\end{equation*}
-Therefore for every $x_0 \in \partial \mc{X}$, it follows from the previous step that
-\begin{equation*}
-\int_{Q_{T_{x_0}}(x_0,c_0(x_0)) \cap \partial_{\varepsilon}\mc{X}} |f(x)|^2 \,dx \leq C(x_0) \varepsilon^{2s} \|f\|_{H^s(\mc{X})}^2,
-\end{equation*}
-where on the right hand side $C(x_0)$ is a constant that may depend on $x_0$, but not on $u$ or $\varepsilon$.
-
-We conclude by taking a collection of cubes that covers $\partial_{\varepsilon}\mc{X}$ for all $\epsilon$ sufficiently small.
-First, we note that by a compactness argument there exists a finite subset of the collection of cubes $\{Q_{T_{x_0}}(x_0,c_0(x_0)/2): x_0 \in \partial\mc{X} \}$ which covers $\partial \mc{X}$, say $Q_{T_{x_1}}(x_1,c_0(x_1)/2),\ldots, Q_{T_{x_N}}(x_N,c_0(x_N)/2)$.
-Then, for any $\varepsilon \leq \min_{i = 1,\ldots,N} c_0(x_i)/2$, it follows from the triangle inequality that
-\begin{equation*}
-\partial_{\varepsilon}\mc{X} \subseteq \bigcup_{i = 1}^{N} Q_{T_{x_i}}(x_i, c_0(x_i)).
-\end{equation*}
-As a result,
-\begin{equation*}
-\int_{\partial_{\varepsilon}\mc{X}} |f(x)|^2 \leq \sum_{i = 1}^{N} \int_{Q_{T_{x_i}}(x_i, c_0(x_i)) \cap \partial_{\varepsilon}(\mc{X})} |f(x)|^2 \leq  \varepsilon^{2s} \|f\|_{H^s(\mc{X})}^2 \sum_{i = 1}^{N}C_0(x_i),
-\end{equation*}
-which proves the claim of~\eqref{pf:approximation_error_nonlocal_laplacian_boundary_0}.
-
-\subsection{Estimate of non-local Sobolev seminorm}
-\label{subsec:estimate_nonlocal_seminorm}
-
-Now, we use the results of the preceding two sections to prove~\eqref{pf:graph_seminorm_ho_2}.
-We will divide our analysis in two cases, depending on whether $s$ is odd or even, but before we do this we state some facts that will be applicable to both cases.
-First, we  recall that $L_{P,\varepsilon}$ is self-adjoint in $L^2(P)$, meaning $\dotp{L_{P,\varepsilon}f}{g}_{P} = \dotp{f}{L_{P,\varepsilon}g}_{P}$ for all $f, g \in L^2(P)$.
-We also recall the definition of the Dirichlet energy $E_{P,\varepsilon}(f;\mc{X})$,
-\begin{equation}
-\label{eqn:dirichlet_energy}
-\dotp{L_{P,\varepsilon}f}{f}_{P} = \frac{1}{\varepsilon^{d + 2}}\int_{\mc{X}} \int_{\mc{X}} \bigl(f(x) - f(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) =: E_{P,\varepsilon}(f;\mc{X}).
-\end{equation}
-Finally, we recall a result of~\cite{green2021}: there exist constants $c_0$ and $C_0$ which do not depend on $M$, such that for all $\varepsilon < c_0$ and for any $f \in H^1(\mc{X};M)$,
-\begin{equation}
-\label{pf:estimate_nonlocal_seminorm_1}
-E_{P,\varepsilon}(f;\mc{X}) \leq C_0 M^2.
-\end{equation}
-\paragraph{Case 1: $s$ odd.}
-Suppose $s$ is odd, so that $s \geq 3$.
-Taking $k = (s - 1)/2$, we use the self-adjointness of $L_{P,\varepsilon}$ to relate the non-local semi-norm $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$ to a non-local Dirichlet energy,
-\begin{equation*}
-\dotp{L_{P,\varepsilon}^sf}{f}_P = \dotp{L_{P,\varepsilon}^{k + 1}f}{L_{P,\varepsilon}^{k}f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\mc{X}).
-\end{equation*}
-We now separate this energy into integrals over $\mc{X}_{k\varepsilon}$ and $\partial_{k\varepsilon}(\mc{X})$,
-\begin{align}
-E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\mc{X}) & = \frac{1}{\varepsilon^{d + 2}}\Biggl\{\int_{\mc{X}_{k\varepsilon}} \int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - L_{P,\varepsilon}^kf(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) \nonumber \\
-& \quad + \int_{\partial_{k\varepsilon}\mc{X}} \int_{\partial_{k\varepsilon}\mc{X}} \bigl(L_{P,\varepsilon}^kf(x) - L_{P,\varepsilon}^kf(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x)\Biggr\} \nonumber \\
-& := E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\mc{X}_{k\varepsilon}) + E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\partial_{k\varepsilon}\mc{X}) \label{pf:estimate_nonlocal_seminorm_1.5}
-\end{align}
-and upper bound each energy separately.
-For the first term, we add and substract $\sigma_{\eta}^k\Delta_P^kf(x)$ and $\sigma_{\eta}^k\Delta_P^kf(x')$ within the integrand, then use the triangle inequality and the symmetry between $x$ and $x'$ to deduce that
-\begin{equation}
-\label{pf:estimate_nonlocal_seminorm_2}
-E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\mc{X}_{k\varepsilon}) \leq 3 \sigma_{\eta}^{2k} E_{P,\varepsilon}(\Delta_P^kf;\mc{X}_{k\varepsilon}) + \frac{2}{\varepsilon^{d + 2}}\int_{\mc{X}_{k\varepsilon}} \int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_P^kf(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x).
-\end{equation}
-Noticing that $\Delta_P^kf \in H^1(\mc{X};\|p\|_{C^{s - 1}(\mc{X})}^kM)$, we use~\eqref{pf:estimate_nonlocal_seminorm_1} to conclude that $E_{P,\varepsilon}(\Delta_P^kf;\mc{X}_{k\varepsilon}) \leq C_0M^2$.
-On the other hand, it follows from Assumption~(K1) and~\eqref{eqn:approximation_error_nonlocal_laplacian_1} that
-\begin{align*}
-\frac{2}{\varepsilon^{d + 2}}\int_{\mc{X}_{k\varepsilon}} \int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_P^kf(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) & \leq \frac{2p_{\max}}{\varepsilon^{2}}\int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_P^kf(x)\bigr)^2 \,dP(x) \\
-& \leq C_1M^2.
-\end{align*}
-Plugging these two bounds into~\eqref{pf:estimate_nonlocal_seminorm_2} gives the desired upper bound on $E_{P,\varepsilon}(L_{P,\varepsilon}^{k};\mc{X}_{k\varepsilon})$.
-
-For the second term in~\eqref{pf:estimate_nonlocal_seminorm_1.5}, we apply Lemmas~\ref{lem:dirichlet_estimate_nonlocal_laplacian} and~\ref{lem:approximation_error_nonlocal_laplacian_boundary} and conclude that,
-\begin{equation*}
-E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\partial_{k\varepsilon}\mc{X}) \leq \frac{4p_{\max}^2}{\varepsilon^2}\|L_{P,\varepsilon}^kf\|_{L^2(\partial_{k\varepsilon}\mc{X})} \leq C M^2.
-\end{equation*}
-\paragraph{Case 2: $s$ even.}
-If $s \in \mathbb{N}$ is even, $s \geq 2$, then letting $k = (s - 2)/2$, the  self-adjointness of $L_{P,\varepsilon}$ implies
-\begin{equation*}
-\dotp{L_{P,\varepsilon}^sf}{f}_P = \|L_{P,\varepsilon}^{k + 1}f\|_P^2.
-\end{equation*}
-As in the first case, we divide the integral up into the interior region $\mc{X}_{k\varepsilon}$ and the boundary region $\partial_{k\varepsilon}\mc{X}$,
-\begin{equation}
-\label{pf:estimate_nonlocal_seminorm_3}
-\|L_{P,\varepsilon}^{k + 1}f\|_P^2 \leq p_{\max} \|L_{P,\varepsilon}^{k + 1}f\|_{L^2(\mc{X})}^2 \leq p_{\max}\biggl\{\int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^{k + 1}f(x)\bigr)^2 \,dP(x) + \int_{\partial_{k\varepsilon}\mc{X}} \bigl(L_{P,\varepsilon}^{k + 1}f(x)\bigr)^2 \,dP(x)\biggr\},
-\end{equation}
-and upper bound each term separately. For the first term, adding and subtracting $\sigma_{\eta}^k \Delta_P^kf(x)$ gives
-\begin{align*}
-\int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^{k + 1}f(x)\bigr)^2 \,dP(x) & \leq 2\int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}\Delta_P^kf(x)\bigr)^2 \,dP(x)  + 2 \int_{\mc{X}_{k\varepsilon}} \Bigl(L_{P,\varepsilon}\bigl(L_{P,\varepsilon}^kf- \sigma_{\eta}\Delta_P^kf\bigr)(x)\Bigr)^2 \,dP(x) \\
-& \overset{(i)}{\leq} CM^2  + 2 \int_{\mc{X}_{k\varepsilon}} \Bigl(L_{P,\varepsilon}\bigl(L_{P,\varepsilon}^kf- \sigma_{\eta}\Delta_P^kf\bigr)(x)\Bigr)^2 \,dP(x) \\
-& \overset{(ii)}{\leq} CM^2  + \frac{Cp_{\max}^2}{\varepsilon^{2}} \|L_{P,\varepsilon}^kf- \sigma_{\eta}\Delta_P^kf\|_{L^2(\mc{X}_{k\varepsilon})}^2 \\
-& \overset{(iii)}{\leq} CM^2,
-\end{align*}
-with $(i)$ following from~\eqref{pf:approximation_error_nonlocal_laplacian_0} since $\Delta_P^kf \in H^2(\mc{X};M\|p\|_{C^{s - 1}(\mc{X})}^l)$, $(ii)$ following from Lemma~\ref{lem:l2estimate_nonlocal_laplacian}, and $(iii)$ following from~\eqref{eqn:approximation_error_nonlocal_laplacian_2}.
-
-Then Lemma~\ref{lem:approximation_error_nonlocal_laplacian_boundary} shows that the second term in~\eqref{pf:estimate_nonlocal_seminorm_3} satisfies
-\begin{equation*}
-\int_{\partial_{k\varepsilon}\mc{X}} \bigl(L_{P,\varepsilon}^{k + 1}f(x)\bigr)^2 \,dP(x) \leq CM^2.
-\end{equation*}
-
-\subsection{Assorted integrals}
-\label{subsec:integrals}
-
-\begin{lemma}
-	\label{lem:l2estimate_nonlocal_laplacian}
-	Assume Model~1.
-Suppose $f \in L^2(U;M)$ for a Borel set $U \subseteq \mc{X}$, and let $L_{P,\varepsilon}$ be defined with respect to a kernel $\eta$ that satisfies~(K1).
-Then there exists a constant $C$ which does not depend on $f$ or $M$ such that
-	\begin{equation}
-	\label{eqn:l2estimate_nonlocal_laplacian}
-	\|L_{P,\varepsilon}f\|_{L^2(U)} \leq \frac{2 p_{\max}}{\varepsilon^2} \|f\|_{L^2(U)}
-	\end{equation}
-\end{lemma}
-
-\begin{lemma}
-	\label{lem:dirichlet_estimate_nonlocal_laplacian}
-	Assume Model~1.
-Suppose $f \in L^2(U;M)$ for a Borel set $U \subseteq \mc{X}$, and let $L_{P,\varepsilon}$ be defined with respect to a kernel $\eta$ that satisfies~(K1).
-Then there exists a constant $C$ which does not depend on $f$ or $M$ such that
-	\begin{equation}
-	\label{eqn:dirichlet_estimate_nonlocal_laplacian}
-	E_{P,\varepsilon}(f;U) \leq \frac{4 p_{\max}^2}{\varepsilon^2} \|f\|_{L^2(U)}^2
-	\end{equation}
-\end{lemma}
-
-\begin{lemma}
-	\label{lem:graph_seminorm_bias2}
-	Assume Model~1.
-Suppose $f \in H^1(\mc{X};M)$, and let $D_if$ be defined with respect to a kernel $\eta$ that satisfies~(K1).
-Then there exists a constant $C$ which does not depend on $f$ or $M$, such that for any $i \in [n]$ and $\bj \in [n]^s$,
-	\begin{equation*}
-	\Ebb\Bigl[|D_{\bj}f(X_i)| \cdot |f(X_i) - f(X_{\bj_1})|\Bigr] \leq C \varepsilon^{2 + dk}M^2,
-	\end{equation*}
-	where $k + 1$ is the number of distinct indices in $i\bj$. 
-\end{lemma}
-
-\paragraph{Proof (of Lemma~\ref{lem:l2estimate_nonlocal_laplacian}).}
-We fix a version of $f \in L^2(U)$, so that we may speak of its pointwise values.
-
-At a given point $x \in U$, we can upper bound $|L_{P,\varepsilon}f(x)|^2$ using the Cauchy-Schwarz inequality as follows,
-\begin{align*}
-|L_{P,\varepsilon}f(x)|^2 & \leq \biggl(\frac{p_{\max}}{\varepsilon^{2 + d}}\biggr)^2 \Biggl(\int_U \bigl(|f(x')| + |f(x)|\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx'\Biggr)^2 \\
-& \leq \biggl(\frac{p_{\max}}{\varepsilon^{2 + d}}\biggr)^2 \Biggl(\int_U \bigl(|f(x')| + |f(x)|\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \cdot \int \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \Biggr) \\
-& = \frac{p_{\max}^2}{\varepsilon^{4 + d}} \int_U \bigl(|f(x')| + |f(x)|\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx'.
-\end{align*}
-The equality follows by the assumption $\int_{\Rd} \eta(\|z\|)\,dx = 1$ in~(K1). Integrating over all $x \in U$, it follows from the triangle inequality that
-\begin{align}
-\|L_{P,\varepsilon}\|_{L^2(U)}^2 & \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx \nonumber \\
-& \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx.
-\label{pf:l2estimate_nonlocal_laplacian_1}
-\end{align}
-Finally, using Fubini's Theorem we determine that
-\begin{equation}
-\int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx = 2 \int_{U} \int_{U} |f(x)|^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx \leq 2 \varepsilon^d \int_{U} |f(x)|^2 \,dx = 2\varepsilon^d \|f\|_{L^2(U)}^2,
-\label{pf:l2estimate_nonlocal_laplacian_2}
-\end{equation}
-and by combining~\eqref{pf:l2estimate_nonlocal_laplacian_1} and~\eqref{pf:l2estimate_nonlocal_laplacian_2} we conclude that
-\begin{equation*}
-\|L_{P,\varepsilon}\|_{L^2(U)}^2 \leq \frac{4p_{\max}^2}{\varepsilon^4} \|f\|_{L^2(U)}^2. 
-\end{equation*}
-
-\paragraph{Proof (of Lemma~\ref{lem:dirichlet_estimate_nonlocal_laplacian}).}
-We have
-\begin{equation*}
-E_{P,\varepsilon}(f) = \frac{1}{\varepsilon^{2 + d}} \int_{U} \int_{U} \bigl(f(x) - f(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) \leq \frac{2p_{\max}^2}{\varepsilon^{2 + d}} \int_{U} \int_{U} \bigl(|f(x)|^2 + |f(x')|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx,
-\end{equation*}
-and the claim follows from~\eqref{pf:l2estimate_nonlocal_laplacian_2}.
-
-\paragraph{Proof (of Lemma~\ref{lem:graph_seminorm_bias2}).}
-Let $G_{n,\varepsilon}[X_{i\bj}]$ be the subgraph induced by vertices $X_i, X_{\bj_1},\ldots,X_{\bj_s}$.
-We make two observations.
-First, in order for $|D_{\bj}f(X_i)| \cdot |f(X_i)  - f(X_j)|$ to be non-zero, it must be the case that the subgraph $G_{n,\varepsilon}[X_{i\bj}]$ is connected.
-Second, noting that for any indices $i$ and $j$,
-\begin{equation*}
-|D_{ij}f(x)| \leq \Bigl(|D_jf(X_i)| + |D_jf(x)|\Bigr)\|\eta\|_{\infty}, 
-\end{equation*}
-a straightforward inductive argument implies that 
-\begin{equation*}
-|D_{\bj}f(X_i)| \leq s\|\eta\|_{\infty}^s \sum_{j \in i\bj} |D_{\bj_s}f(X_j)|.
-\end{equation*}
-Combining these two observations, we reduce the task to upper bounding the product of two (first-order) differences,
-\begin{align*}
-\mathbb{E}\Bigl[ |D_{\bj}f(X_i)| |f(X_i) - f(X_{\bj_1})| \Bigr] & = \mathbb{E}\Bigl[ |D_{\bj}f(X_i)| |f(X_i) - f(X_{\bj_1})| \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr] \\
-& \leq s\|\eta\|_{\infty}^s \sum_{j \in i\bj} \mathbb{E}\Bigl[ |D_{\bj_s}f(X_j)| \cdot |f(X_i) - f(X_{\bj_1})| \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\}   \Bigr] \\
-& \leq s\|\eta\|_{\infty}^s \sum_{j \in i\bj} \mathbb{E}\Bigl[ |f(X_j) - f(X_{\bj_s})| \cdot |f(X_i) - f(X_{\bj_1})| \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\}   \Bigr]
-\end{align*}
-Next, from the Cauchy-Schwarz inequality we have that for any $j \in \bj$,
-\begin{align*}
-& \mathbb{E}\Bigl[ |f(X_j) - f(X_{\bj_s})| \cdot |f(X_i) - f(X_{\bj_1})| \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\}   \Bigr] \\
-& \quad \leq \sqrt{\mathbb{E}\Bigl[ |f(X_j) - f(X_{\bj_s})|^2 \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr]} \cdot \sqrt{\mathbb{E}\Bigl[ |f(X_j) - f(X_{\bj_s})|^2 \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr]} \\
-& \quad = \mathbb{E}\Bigl[ |f(X_j) - f(X_{i})|^2 \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr],
-\end{align*}
-with the equality following since each $X_i$ are identically distributed. Marginalizing out the contribution of all indices in $\bj$ not equal to $i$ or $j$ gives
-\begin{align}
-\mathbb{E}\Bigl[ |f(X_j) - f(X_{i})|^2 \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr] & \leq \bigl((s + 1)p_{\max}\nu_d\varepsilon^d\bigr)^{|i\bj\setminus\{j \cup i\}|} \cdot \mathbb{E}\Bigl[ |f(X_j) - f(X_{i})|^2 \1\{\|X_i - X_j\| \leq \varepsilon\}\Bigr] \nonumber \\
-& \leq \bigl((s + 1)p_{\max}\nu_d\varepsilon^d\bigr)^{|i\bj\setminus\{j \cup i\}|} \cdot p_{\max}^2 \nu_d \varepsilon^{2 + d} M^2 \label{pf:graph_seminorm_bias2_1}
-\end{align}
-with the second inequality following from the proof of Lemma~1 in~\cite{green2021}.
-Finally, we notice that $|i\bj\setminus \{i \cup j\}| + 1 = k$, so that~\eqref{pf:graph_seminorm_bias2_1} gives the desired result.
-
-\section{Graph Sobolev semi-norm, manifold domain}
-\label{sec:graph_quadratic_form_manifold}
-In this section we prove Proposition~7.
-Note that when $s = 1$, the desired upper bound (Equation~(42) in the main text) follows immediately from Lemma~\ref{lem:dirichlet_energy_sobolev} and Markov's inequality.
-
-On the other hand when $s = 2$ or $s = 3$, we prove Proposition~7 by first establishing some intermediate results, many of which are analogous to results we have already shown in the flat Euclidean case.
-Indeed, in some ways the proof will be simpler in the manifold setting than in the flat Euclidean case: there is no boundary, and we do not need to analyze the iterated nonlocal Laplacian $L_{P,\varepsilon}^j$ for $j > 1$.
-
-That being said, as mentioned in our main text, in the manifold setting there is some extra error induced by using Euclidean rather than geodesic distance.
-We upper bound this error by comparing $L_{P,\varepsilon}$ to an alternative nonlocal Laplacian $\wt{L}_{P,\varepsilon}$, which is defined with respect to geodesic distance.
-Precisely, let $d_{\mc{X}}(x,x')$ denote the geodesic distance between $x,x' \in \mc{X}$, and define
-\begin{equation*}
-\wt{L}_{P,\varepsilon}f(x) := \int_{\mc{X}} \bigl(f(x') - f(x)\bigr) \eta \biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr) p(x') \,dx'.
-\end{equation*}
-
-We show the following results, each of which hold under the same assumptions as Proposition~7.
-\begin{itemize}
-	\item In Section~\ref{subsec:manifold_decomposition_graph_seminorm} we show that the graph Sobolev seminorm $\dotp{L_{n,\varepsilon}^sf}{f}_n$ is upper bounded by the sum of a nonlocal seminorm and a pure bias term: specifically, with probability at least $1 - 2\delta$,
-	\begin{equation}
-	\label{pf:graph_seminorm_manifold_1}
-	\dotp{L_{n,\varepsilon}^sf}{f}_n \leq \frac{\dotp{L_{P,\varepsilon}^sf}{f}_P}{\delta} + C_1\frac{\varepsilon^2}{n\varepsilon^{2s + m}}M^2.
-	\end{equation}
-	This upper bound is essentially the same as~\eqref{pf:graph_seminorm_ho_1}, but with the intrinsic dimension $m$ taking the place of the ambient dimension $d$.
-The pure bias term will be of at most constant order when $\varepsilon \gtrsim n^{-1/(2(s-1) + m)}$.
-	\item In Section~\ref{subsec:error_euclidean_distance}, we show that the error incurred by using the ``wrong'' metric is negligible. Precisely, we find that
-	\begin{equation}
-	\label{eqn:nonlocal_laplacian_geodesic_error}
-	\|L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f\|_{L^2(\mc{X})}^2 \leq C_2 \varepsilon^2 |f|_{H^1(\mc{X})}^2.
-	\end{equation}
-	\item In Section~\ref{subsec:manifold_approximation_error_nonlocal_laplacian}, we analyze the approximation error of $\wt{L}_{P,\varepsilon}$.
-We show that when $f \in H^2(\mc{X})$ and $p \in C^1(\mc{X})$,
-	\begin{equation}
-	\label{eqn:nonlocal_laplacian_approximation_error_manifold_l2}
-	\|\wt{L}_{P,\varepsilon}f\|_{L^2(\mc{X})}^2 \leq C_3 \|f\|_{H^2(\mc{X})}^2,
-	\end{equation}
-	whereas if $f \in H^3(\mc{X})$ and $p \in C^2(\mc{X})$, 
-	\begin{equation}
-	\label{eqn:nonlocal_laplacian_approximation_error_manifold_sobolev}
-	\|\wt{L}_{P,\varepsilon}f - \sigma_{\eta}\Delta_Pf\|_{L^2(\mc{X})}^2 \leq C_3 \varepsilon^2 \|f\|_{H^3(\mc{X})}^2.
-	\end{equation}
-	\item In Section~\ref{subsec:manifold_estimate_nonlocal_seminorm}, we use the results of the preceding two sections to show that if $f \in H^s(\mc{X})$ and $p \in C^{s - 1}(\mc{X})$, then 
-	\begin{equation}
-	\label{eqn:manifold_nonlocal_seminorm}
-	\dotp{L_{P,\varepsilon}^sf}{f}_P \leq C_4\|f\|_{H^s(\mc{X})}^2.
-	\end{equation}
-	\item In Section~\ref{subsec:manifold_integrals} we state some technical results used in the previous sections.
-\end{itemize}
-We point out that when $f$ is H\"{o}lder smooth, results analogous to~\eqref{eqn:nonlocal_laplacian_approximation_error_manifold_sobolev} have been established in \citet{calder2019}.
-When $f$ is Sobolev smooth, our analysis (which relies heavily on Taylor expansions) is largely similar, except that the remainder term in the relevant Taylor expansion will be bounded in $L^2(\mc{X})$ norm rather than $L^{\infty}(\mc{X})$ norm.
-This is analogous to the situation in the flat Euclidean model.
-
-In the proof of~\eqref{pf:graph_seminorm_manifold_1}-\eqref{eqn:manifold_nonlocal_seminorm}, we recall the following estimates from differential geometry: (i) letting $K_0$ be an upper bound on the absolute value of the sectional curvatures of $\mc{X}$, $K_0 \leq 2R$, and  letting (ii) $i_0$ be a lower bound on the injectivity radius of $\mc{X}$, $i_0 \geq \pi R$; see Proposition~1 of \cite{aamari2019}.
-Additionally, recall that for all $\delta < i_0$, the exponential map $\exp_x: B_m(0,\delta) \subset T_x(\mc{X}) \to B_{\mc{X}}(x,\delta) \subset \mc{X}$ is a diffeomorphism for all $x \in \mc{X}$.
-We shall therefore always assume $\varepsilon < i_0$.
-
-\paragraph{Proof (of Proposition~7).} Follows immediately from~\eqref{pf:graph_seminorm_manifold_1} and~\eqref{eqn:manifold_nonlocal_seminorm}. \qed
-
-\subsection{Decomposition of graph Sobolev seminorm}
-\label{subsec:manifold_decomposition_graph_seminorm}
-The proof of~\eqref{pf:graph_seminorm_manifold_1} is identical to the proof of~\eqref{pf:graph_seminorm_ho_1}, except substituting the intrinsic dimension $m$ for ambient dimension $d$, and using Lemma~\ref{lem:manifold_graph_seminorm_bias2} rather than Lemma~\ref{lem:graph_seminorm_bias2}.
-
-\subsection{Error due to Euclidean Distance}
-\label{subsec:error_euclidean_distance}
-
-In this section, we prove~\eqref{eqn:nonlocal_laplacian_geodesic_error}. By applying Cauchy-Schwarz we obtain an upper bound on $|L_{P,\varepsilon}f(x) - \wt{L}_{P,\varepsilon}f(x)|^2$:
-\begin{align}
-\bigl[L_{P,\varepsilon}f(x) - \wt{L}_{P,\varepsilon}f(x)\bigr]^2 & \leq \frac{p_{\max}^2}{\varepsilon^{2(2 + m)}} \int_{\mc{X}} \bigl[f(x') - f(x)\bigr]^2 \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') \nonumber \\
-& \quad \cdot \int_{\mc{X}} \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') \nonumber \\
-& = \frac{1}{\varepsilon^{2(2 + m)}} A_1(x) \cdot A_2(x) \label{pf:error_euclidean_distance_0}
-\end{align} 
-Thus we have upper bounded $|L_{P,\varepsilon}f(x) - \wt{L}_{P,\varepsilon}f(x)|^2$ by the product of two terms, each of which we now suitably bound.
-To do so, we will use the following estimate, from Proposition 4 of \cite{trillos2019}: for all $\|x' - x\| \leq R/2$,
-\begin{equation}
-\label{eqn:distance_error}
-\|x' - x\| \leq d_{\mc{X}}(x',x) \leq \|x' - x\| + \frac{8}{R^2} \|x' - x\|^3.
-\end{equation}
-From here forward we will assume $\varepsilon < R/2$. 
-\paragraph{Upper bound on $A_1(x)$.}
-Consequently $\eta(\|x' - x\|/\varepsilon) \geq \eta(d_{\mc{X}}(x',x)/\varepsilon)$.
-Furthermore, letting $L_{\eta}$ denote the Lipschitz constant of $\eta$, and setting $\wt{\varepsilon} := (1 + 27\varepsilon^2/R^2)\varepsilon$ we have that
-\begin{equation*}
-\biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \leq \frac{L_{\eta} 8 \varepsilon^2}{R^2} \cdot \1\bigl\{d_{\mc{X}}(x',x) \leq \varepsilon\bigr\} + \|\eta\|_{\infty} \cdot 1\{\varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon}\}.
-\end{equation*}
-Thus,
-\begin{equation*}
-A_1(x) \leq \frac{8L_{\eta}\varepsilon^2}{R^2}\int_{\mc{X}}\bigl[f(x') - f(x)\bigr]^2 \1\{\|x' - x\| \leq \varepsilon\} \,d\mu(x') + \|\eta\|_{\infty} \int_{\mc{X}}\bigl[f(x') - f(x)\bigr]^2 \1\bigl\{\varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon}\bigr\} \,d\mu(x') \\
-\end{equation*}
-Integrating over $\mc{X}$, we conclude from Lemma~\ref{lem:dirichlet_energy_remainder} and Lemma~3.3 of \citep{burago2014} and  that
-\begin{equation*}
-\int_{\mc{X}} A_1(x) \,d\mu(x) \leq \frac{8L_{\eta}\nu_m\varepsilon^2}{R^2(m + 2)} \Bigl(1 + CmK_0R^2\Bigr) \varepsilon^{m + 2} |f|_{H^1(\mc{X})}^2 + C\|\eta\|_{\infty}\varepsilon^{m + 4} |f|_{H^1(\mc{X})}^2 =: C_5 \varepsilon^{m + 4}|f|_{H^1(\mc{X})}.
-\end{equation*}
-
-\paragraph{Upper bound on $A_2(x)$.}
-Integrating over $x' \in \mc{X}$, we see that
-\begin{align}
-\int_{\mc{X}} \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') & \leq \frac{8L_{\eta}\varepsilon^2}{R^2} \int_{\mc{X}} \1\bigl\{d_{\mc{X}}(x',x)\bigr\} \,d\mu(x') + p_{\max}\|\eta\|_{\infty} \int_{\mc{X}} \1\bigl\{ \varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon} \bigr\} \,d\mu(x') \nonumber \\
-& = \frac{8L_{\eta}\varepsilon^2}{R^2} \cdot \mu\bigl(B(x,\varepsilon)\bigr) +  p_{\max}\|\eta\|_{\infty} \Bigl[\mu\bigl(B(x,\wt{\varepsilon})\bigr) - \mu\bigl(B(x,\varepsilon)\bigr) \Bigr].
-\label{pf:error_euclidean_distance_1}
-\end{align}
-Equation (1.36) in \cite{trillos2019} states that
-\begin{equation*}
-\bigl|\mu(B_\mc{X}(x,\varepsilon))  - \omega_m \varepsilon^m\bigr|  \leq CmK_0\varepsilon^{m + 2},
-\end{equation*}
-where we recall $K_0$ is an upper bound on the sectional curvature of $\mc{X}$. Plugging this back into~\eqref{pf:error_euclidean_distance_1}, we conclude that
-\begin{align*}
-\int_{\mc{X}} \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') & \leq  \frac{8L_{\eta}\varepsilon^2}{R^2}\Bigl[\omega_m\varepsilon^{m} + CmK_0\varepsilon^{m + 2}\Bigr] + \|\eta\|_{\infty} \Bigl[\omega_m(\wt{\varepsilon}^m - {\varepsilon}^m) + 2CmK_0\varepsilon^{m + 2}\Bigr] \\
-& \leq \frac{8L_{\eta}\varepsilon^2}{R^2}\Bigl[\omega_m\varepsilon^{m} + R^2CmK_0\varepsilon^{m}\Bigr] + \|\eta\|_{\infty} \varepsilon^{m + 2}\Bigl[\frac{27\omega_m}{R^2} + 2CmK_0\Bigr] \\
-& =: C_6 \varepsilon^{m + 2}.
-\end{align*}
-
-\paragraph{Putting together the pieces.}
-Plugging our upper bounds on $A_1(x)$ and $A_2(x)$ back into~\eqref{pf:error_euclidean_distance_0}, we deduce that
-\begin{align*}
-\|\wt{L}_{P,\varepsilon}f - L_{P,\varepsilon}f\|_{L^2(\mc{X})}^2 & \leq \frac{1}{\varepsilon^{2(2 + m)}} \int_{\mc{X}} A_1(x) \cdot A_2(x) \,d\mu(x) \\
-& \leq \frac{C_6}{\varepsilon^{(2 + m)}} \int_{\mc{X}} A_1(x) \,d\mu(x) \\
-& \leq C_5 C_6 \varepsilon^2 |f|_{H^1(\mc{X})}^2,
-\end{align*}
-thus proving the claimed result.
-
-\subsection{Approximation Error of non-local Laplacian}
-\label{subsec:manifold_approximation_error_nonlocal_laplacian}
-Fix $x \in \mc{X}$.
-We begin with a pointwise estimate of $\wt{L}_{P,\varepsilon}f$, facilitated by expressing $w(v) = f(\exp_x(v))$ and $q(v) = p(\exp_x(v))$ in normal coordinates, as in \citep{calder2019}.
-Let $J_x(\cdot)$ be the Jacobian of the exponential map $\exp_x$, we have
-\begin{align*}
-\wt{L}_{P,\varepsilon}f(x) & = \frac{1}{\varepsilon^{m + 2}} \int_{\mc{X}} \bigl(f(x') - f(x)\bigr) \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr) \,dP(x') \\
-& = \frac{1}{\varepsilon^{m + 2}} \int_{B(0,\varepsilon) \subset T_x(\mc{X})} \bigl(w(v) - w(0)\bigr) \eta\biggl(\frac{\|v\|}{\varepsilon}\biggr) J_x(v) q(v) \,dv \\
-& = \frac{1}{\varepsilon^{2}}\biggl\{\int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr) \eta(\|v\|) q(\varepsilon v) \,dv + \int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr) \eta(\|v\|) q(\varepsilon v) \bigl(J_x(\varepsilon v) - 1\bigr) \,dv \biggr\} \\
-& = A_1(x) + A_2(x)
-\end{align*}
-Note that $w$ and $q$ have the same smoothness properties as $f$ and $p$.
-Moreover, arguing exactly as we did in the flat Euclidean case, we can show that when $f \in H^2(\mc{X})$ and $p \in C^1(\mc{X})$, then
-\begin{equation*}
-\|A_1\|_{L^2(\mc{X})}^2 \leq C \|f\|_{H^2(\mc{X})}^2
-\end{equation*}
-whereas if $f \in H^3(\mc{X})$ and $p \in C^2(\mc{X})$ then 
-\begin{equation*}
-\|A_1 - \sigma_{\eta} \Delta_Pf\|_{L^2(\mc{X})}^2 \leq C\|f\|_{H^3(\mc{X})}^2 \varepsilon^2.
-\end{equation*}
-
-Therefore it remains only to upper bound $A_2$ in $L^2(\mc{X})$ norm.
-To do so, we recall (1.34) of \cite{trillos2019}: for any $\varepsilon < i_0$ and all $x \in \mc{X}$, the Jacobian $J_x(v)$ satisfies the upper bound
-\begin{equation*}
-|J_x(v) - 1| \leq CmK_0\varepsilon^2, \quad  \textrm{for all} ~~ v \in B(0,\varepsilon) \subseteq T_x(\mc{X}).
-\end{equation*}
-Combining this estimate with the Cauchy-Schwarz inequality, we conclude that
-\begin{align*}
-\|A_2\|_{L^2(\mc{X})}^2 & \leq Cm^2K_0^2 \biggl[\int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr)^2 \eta(\|v\|) q(\varepsilon v) \,dv\biggr] \cdot \biggl[\int_{B(0,1)} \eta(\|v\|) q(\varepsilon v) \,dv\biggr] \\
-& \leq Cm^2K_0^2 \sigma_{\eta} (1 + L_q\varepsilon)  \int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr)^2 \eta(\|v\|) q(\varepsilon v) \,dv \\
-& \leq Cm^2K_0^2 \sigma_{\eta}^2 (1 + L_q\varepsilon) p_{\max}  \varepsilon^2 |f|_{H^1(\mc{X})}^2,
-\end{align*}
-with the final inequality following from (3.2) of~\cite{burago2014}. Combining our estimates on $A_1$ and $A_2$ yields the claim.
-
-\subsection{Estimate of non-local Sobolev seminorm}
-\label{subsec:manifold_estimate_nonlocal_seminorm}
-In this subsection we establish that the upper bound~\eqref{eqn:manifold_nonlocal_seminorm} holds when $f \in H^s(\mc{X})$ and $p \in C^{s - 1}(\mc{X})$. We first consider $s = 2$, and then $s = 3$.
-
-\paragraph{Case 1: $s = 2$.}
-When $s = 2$, the triangle inequality implies that
-\begin{equation*}
-\dotp{L_{P,\varepsilon}^sf}{f}_P \leq 2p_{\max}\Bigl(\|L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}\|_{L^2(\mc{X})}^2 + \|\wt{L}_{P,\varepsilon}f\|_{L^2(\mc{X})}^2\Bigr)
-\end{equation*}
-The first term on the right hand side is upper bounded in~\eqref{eqn:nonlocal_laplacian_geodesic_error}, and the second term is upper bounded in~\eqref{eqn:nonlocal_laplacian_approximation_error_manifold_l2}.
-Together these estimates imply the claim.
-
-\paragraph{Case 2: $s = 3$.}
-When $s = 3$, the triangle inequality implies that
-\begin{equation*}
-\dotp{L_{P,\varepsilon}^sf}{f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}f;\mc{X}) \leq 3\Bigl( E_{P,\varepsilon}(L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f;\mc{X}) +  E_{P,\varepsilon}(\wt{L}_{P,\varepsilon}f - \sigma_{\eta} \Delta_Pf;\mc{X}) + \sigma_{\eta}^2 E_{P,\varepsilon}(\Delta_Pf;\mc{X})\Bigr)
-\end{equation*}
-We now upper bound each of the three terms on the right hand side of the above inequality.
-First, we note that by Lemma~\ref{lem:dirichlet_energy_l2} and~\eqref{eqn:nonlocal_laplacian_geodesic_error},
-\begin{equation*}
-E_{P,\varepsilon}(L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f;\mc{X}) \leq  \frac{C}{\varepsilon^2}\|L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f\|_{L^2(\mc{X})}^2 \leq C |f|_{H^1(\mc{X})}^2.
-\end{equation*}
-An equivalent upper bound on $E_{P,\varepsilon}(\wt{L}_{P,\varepsilon}f - \sigma_{\eta} \Delta_Pf;\mc{X})$ follows from Lemma~\ref{lem:dirichlet_energy_l2} and~\eqref{eqn:nonlocal_laplacian_approximation_error_manifold_sobolev}.
-Finally, we notice that $f \in H^3(\mc{X})$ and $p \in C^2(\mc{X})$ implies $\Delta_Pf \in H^1(\mc{X})$, and furthermore $|\Delta_Pf|_{H^1(\mc{X})} \leq \|p\|_{C^2(\mc{X})} \cdot \|f\|_{H^3(\mc{X})}$.
-We conclude from Lemma~\ref{lem:dirichlet_energy_sobolev} that
-\begin{equation*}
-E_{P,\varepsilon}(\Delta_Pf;\mc{X}) \leq C |\Delta_Pf|_{H^1(\mc{X})}^2 \leq C \|f\|_{H^3(\mc{X})}^2,
-\end{equation*}
-where in the final inequality we have absorbed $\|p\|_{C^2(\mc{X})}$ into the constant $C$. Together, these upper bounds prove the claim.
-
-\subsection{Integrals}
-\label{subsec:manifold_integrals}
-Recall the Dirichlet energy $E_{P,\varepsilon}(f;\mc{X}) = \dotp{L_{P,\varepsilon}f}{f}_P$, defined in~\eqref{eqn:dirichlet_energy}.
-Now we establish some estimates on $E_{P,\varepsilon}(f;\mc{X})$ under Model~2, and under various assumptions regarding the regularity of $f$.
-\begin{lemma}
-	\label{lem:dirichlet_energy_l2}
-	Suppose Model~2, and additionally that $f \in L^2(\mc{X})$. Then there exists a constant $C$ such that
-	\begin{equation}
-	\label{eqn:dirichlet_energy_l2}
-	E_{P,\varepsilon}(f;\mc{X}) \leq \frac{C}{\varepsilon^2} \|f\|_{L^2(\mc{X})}^2.
-	\end{equation}
-\end{lemma}
-\begin{lemma}
-	\label{lem:dirichlet_energy_sobolev}
-	Suppose Model~2, and additionally that $f \in H^1(\mc{X})$. Then there exist constants $c$ and $C$ which do not depend on $f$ such that for any $0 < \varepsilon < c$,
-	\begin{equation}
-	\label{eqn:dirichlet_energy_sobolev}
-	E_{P,\varepsilon}(f;\mc{X}) \leq C |f|_{H^1(\mc{X})}^2.
-	\end{equation}
-\end{lemma}
-
-We use Lemma~\ref{lem:dirichlet_energy_remainder} to help upper bound the error incurred by using $\|\cdot\|$ rather than $d_{\mc{X}}(\cdot,\cdot)$.
-Recall the notation $\wt{\varepsilon} = (1 + 27\varepsilon^2/R^2)\varepsilon$, where $R$ is the reach of $\mc{X}$.
-\begin{lemma}
-	\label{lem:dirichlet_energy_remainder}
-	Suppose Model~2, and additionally that $f \in H^1(\mc{X})$. There exist constants $c$ and $C$ such that for any $\varepsilon < c$,
-	\begin{equation}
-	\label{eqn:dirichlet_energy_remainder}
-	\int_{\mc{X}} \int_{\mc{X}} \bigl(f(x') - f(x)\bigr)^2 \1\{\varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon}\} \,d\mu(x') \,d\mu(x) \leq C \varepsilon^{4 + m} \|f\|_{H^1(\mc{X})}^2
-	\end{equation}
-\end{lemma}
-
-Finally, we use Lemma~\ref{lem:manifold_graph_seminorm_bias2} to show that the pure bias component of $\dotp{L_n^sf,f}_n$ is small in expectation.
-This is analogous to Lemma~\ref{lem:graph_seminorm_bias2}, except assuming Model~2 rather than Model~1.
-\begin{lemma}
-	\label{lem:manifold_graph_seminorm_bias2}
-	Assume Model~2.
-Suppose $f \in H^1(\mc{X})$, and let $D_if$ be defined with respect to a kernel $\eta$ that satisfies~(P5).
-Then there exists a constant $C$ which does not depend on $f$ or $n$, such that for any $i \in [n]$ and $\bj \in [n]^s$,
-	\begin{equation*}
-	\Ebb\Bigl[|D_{\bj}f(X_i)| \cdot |f(X_i) - f(X_{\bj_1})|\Bigr] \leq C \varepsilon^{2 + mk} \cdot \|f\|_{H^1(\mc{X})}^2,
-	\end{equation*}
-	where $k + 1$ is the number of distinct indices in $i\bj$. 
-\end{lemma}
-
-\paragraph{Proof (of Lemmas~\ref{lem:dirichlet_energy_l2} and~\ref{lem:dirichlet_energy_sobolev}).}
-Define the non-local energy $\wt{E}_{P,\varepsilon}$ with respect to geodesic distance,
-\begin{equation*}
-\wt{E}_{P,\varepsilon}(f;{\mc{X}}) := \dotp{\wt{L}_{P,\varepsilon}f}{f}_{P} = \int_{\mc{X}} \int_{\mc{X}} \bigl(f(x') - f(x)\bigr)^2 \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr) \,dP(x') \,dP(x).
-\end{equation*}
-From the lower bound in~\eqref{eqn:distance_error}, it follows that $E_{P,\varepsilon}(f;X) \leq \wt{E}_{P,\varepsilon}(f;{\mc{X}})$, and from the upper bounds $p(x) \leq p_{\max}$ and $\eta(|x|) \leq \|\eta\|_{\infty} \cdot \1\{x \in [-1,1]\}$ we further have
-\begin{equation*}
-\wt{E}_{P,\varepsilon}(f;{\mc{X}}) \leq p_{\max}^2 \|\eta\|_{\infty} \cdot \int_{\mc{X}} \int_{B_{\mc{X}}(\varepsilon)} \bigl(f(x') - f(x)\bigr)^2 \,d\mu(x') \,d\mu(x).
-\end{equation*}
-The estimates~\eqref{eqn:dirichlet_energy_l2} and~\eqref{eqn:dirichlet_energy_sobolev} then respectively follow from (3.1) and Lemma~3.3 of \cite{burago2014}.
-
-\paragraph{Proof (of Lemma~\ref{lem:dirichlet_energy_remainder}).}
-Following exactly the steps of the proof of Lemma~3.3 of \citet{burago2014}, but replacing all references to a ball of radius $r$ by references to the set difference between balls of radius $\wt{\varepsilon}$ and $\varepsilon$, we obtain that
-\begin{equation*}
-\int_{\mc{X}} \int_{\mc{X}} \bigl(f(x') - f(x)\bigr)^2 \1\{\varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon}\} \,d\mu(x') \,d\mu(x) \leq (1 + CmK_0\varepsilon^2) \cdot \int_{\mc{X}} \int_{B_{m}(0,\wt{\varepsilon})} |d_x^{1}f(v)|^2 \,dv \,d\mu(x).
-\end{equation*}
-From (2.7) of~\citet{burago2014}, we further have
-\begin{equation*}
-\int_{\mc{X}} \int_{B_{m}(0,\wt{\varepsilon})} |d_x^{1}f(v)|^2 \,dv \,d\mu(x)  = \frac{\nu_m}{2 + m} (\wt{\varepsilon}^{2 + m} - \varepsilon^{2 + m}) \int_{\mc{X}} |d_x^1f|^2 \,d\mu(x) = 27\frac{\nu_m}{(2 + m)R^2} \varepsilon^{4 + m} \|d^1f\|_{L^2(\mc{X})}^2.
-\end{equation*}
-Recalling that $\|d^1f\|_{L^2(\mc{X})}^2 \leq \|f\|_{H^1(\mc{X})}^2$, we see that this implies the claim of Lemma~\ref{lem:dirichlet_energy_remainder}.
-
-\paragraph{Proof (of Lemma~\ref{lem:manifold_graph_seminorm_bias2}).}
-The proof of Lemma~\ref{lem:manifold_graph_seminorm_bias2} is identical to the proof of Lemma~\ref{lem:graph_seminorm_bias2}, upon substituting the ambient dimension $m$ for the intrinsic dimension $d$, and using Lemma~\ref{lem:dirichlet_energy_sobolev} rather than Lemma~\ref{lem:dirichlet_estimate_nonlocal_laplacian} to establish~\eqref{pf:graph_seminorm_bias2_1}.
-
-\section{Lower bound on empirical norm}
-\label{sec:empirical_norm}
-In this Section we prove Proposition~6 (in Section~\ref{subsec:empirical_norm_sobolev}).
-We also prove an analogous result when $\mc{X}$ is a manifold as in Model~2 (in Section~\ref{subsec:empirical_norm_sobolev_manifold}).
-
-\subsection{Proof of Proposition~6}
-\label{subsec:empirical_norm_sobolev}
-In this section we establish Proposition~6.
-As mentioned, the proof of this Proposition follows from the Gagliardo-Nirenberg interpolation inequality, and a one-sided Bernstein's inequality (Lemma~\ref{lem:one_sided_bernstein}).
-
-\begin{lemma}[Gagliardo-Nirenberg interpolation inequality]
-	\label{lem:gagliardo_nirenberg}
-	Suppose Model~1, and that $f \in H^s(\mc{X})$ for some $s \geq d/4$. Then there exist constants $C_1$ and $C_2$ that do not depend on $f$, such that
-	\begin{equation}
-	\label{eqn:gagliardo_nirenberg}
-	\|f\|_{L^4(\mc{X})} \leq C_1 |f|_{H^s(\mc{X})}^{d/4s} \|f\|_{L^2(\mc{X})}^{1 - d/(4s)} + C_2 \|f\|_{L^2(\mc{X})}
-	\end{equation}
-\end{lemma}
-
-\paragraph{Proof (of Proposition~6).}
-Rearranging~\eqref{eqn:gagliardo_nirenberg} and raising both sides to the $4$th power, we see that
-\begin{equation*}
-\frac{\Ebb[f^4(X)]}{\|f\|_P^4} \leq C \biggl(\frac{\|f\|_{L^4(\mc{X})}}{\|f\|_{L^2(\mc{X})}}\biggr)^4 \leq C_1\biggl(\frac{|f|_{H^s(\mc{X})}}{\|f\|_{L^2(\mc{X})}}\biggr)^{d/s} + C_2,
-\end{equation*}
-here the constants $C_1,C_2$ are not the same as in~\eqref{eqn:gagliardo_nirenberg}.
-Therefore taking the constant $C$ in the statement of Proposition~6 to be sufficiently large relative to $C_1$ and $C_2$, we have that
-\begin{equation*}
-C_1\biggl(\frac{|f|_{H^s(\mc{X})}}{\|f\|_{L^2(\mc{X})}}\biggr)^{d/s} \leq \frac{\delta n}{64},
-\end{equation*} 
-and consequently 
-\begin{equation*}
-\frac{\Ebb[f^4(X)]}{\|f\|_P^4} \leq \frac{\delta n}{8} + 8C_2^3.
-\end{equation*}
-The claim then follows from Lemma~\ref{lem:one_sided_bernstein}, upon taking $c = 1/(64C_2^3)$ in the statement of Proposition~6.
-
-\subsection{Proof of Proposition~9}
-\label{subsec:empirical_norm_sobolev_manifold}
-The proof of Proposition~9 follows exactly the same steps as the proof of Proposition~6, upon replacing Lemma~\ref{lem:gagliardo_nirenberg} by Lemma~\ref{lem:gagliardo_nirenberg_manifold}.
-\begin{lemma}[(c.f Theorem~3.70 of~\citet{aubin2012})]
-	\label{lem:gagliardo_nirenberg_manifold}
-	Suppose Model~2, and that $f \in H^s(\mc{X})$ for some $s \geq m/4$. Then there exist constants $C_1$ and $C_2$ that do not depend on $f$, such that
-	\begin{equation}
-	\label{eqn:gagliardo_nirenberg_manifold}
-	\|f\|_{L^4(\mc{X})} \leq C_1 |f|_{H^s(\mc{X})}^{m/4s} \|f\|_{L^2(\mc{X})}^{1 - m/(4s)} + C_2 \|f\|_{L^2(\mc{X})}.
-	\end{equation}
-\end{lemma}
-
-\section{Proof of Main Results}
-\label{sec:proofs_main_results}
-
-\subsection{Estimation Results}
-
-\paragraph{Proof of Theorem~1.}
-We condition on the event that the design points $X_1,\ldots,X_n$ satisfy
-\begin{equation}
-\label{pf:laplacian_eigenmaps_estimation_fo_1}
-\dotp{L_{n,\varepsilon}f_0}{f_0}_n \leq \frac{C}{\delta}M^2 \quad \textrm{and} \quad \lambda_k \geq \min\{\lambda_k(\Delta_P), \varepsilon^{-2}\}~~\textrm{for all $2 \leq k \leq n$.}
-\end{equation}
-Note that by Propositions~3 and~4, these statements are both satisfied with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\}$.
-
-Conditional on~\eqref{pf:laplacian_eigenmaps_estimation_fo_1}, we have from Lemma~\ref{lem:fixed_graph_estimation} that for any $0 \leq K \leq n$,
-\begin{equation*}
-\|\wh{f} - f_0\|_n^2 \leq C\biggl\{\frac{M^2}{\delta (\lambda_{K + 1}(\Delta_P) \wedge \varepsilon^{-2})} + \frac{K}{n}\biggr\},
-\end{equation*}
-either deterministically (when $K = 0$), or with probability at least $1 - \exp(-K)$ (when $K \geq 1$).
-Further, from the bounds $\varepsilon \leq c_0 K^{-1/d}$ (Assumption~(P1)) and $\lambda_{K + 1}(\Delta_P) \geq c (K + 1)^{2/d}$ (Weyl's Law) we can simply the above expression to the following,
-\begin{equation}
-\label{pf:laplacian_eigenmaps_estimation_fo_2}
-\|\wh{f} - f_0\|_n^2 \leq C\biggl\{\frac{M^2}{\delta}(K + 1)^{-2/d} + \frac{K}{n}\biggr\}.
-\end{equation}
-We now upper bound the right hand side of~\eqref{pf:laplacian_eigenmaps_estimation_fo_2}, based on the value of $K$ chosen in~(P1).
-When possible we choose $K = \floor{M^2n}^{d/(2 + d)}$ to balance bias and variance, in which case~\eqref{pf:laplacian_eigenmaps_estimation_fo_2} implies
-\begin{equation*}
-\|\wh{f} - f_0\|_n^2 \leq \frac{C}{\delta} M^2 (M^2n)^{-2/(2 + d)}.
-\end{equation*}
-If $M^2 < n^{-1}$, then we take $K = 1$, and from~\eqref{pf:laplacian_eigenmaps_estimation_fo_2} we get
-\begin{equation*}
-\|\wh{f} - f_0\|_n^2 \leq \frac{C}{n\delta}.
-\end{equation*}
-Finally if $M > n^{1/d}$, we take $K = n$. In this case, we note that $\wh{f}(X_i) = Y_i$ for all $i = 1,\ldots,n$, and it immediately follows that
-\begin{equation*}
-\|\wh{f} - f_0\|_n^2 = \frac{1}{n}\sum_{i = 1}^{n} w_i^2 \leq 5,
-\end{equation*}
-with probability at least $1 - \exp(-n)$. Combining these three separate cases yields the conclusion of Theorem~1.
-
-\paragraph{Proof of Theorem~3.}
-Follows identically to the proof of Theorem~1, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~4 rather than Proposition~3 and Assumption~(P3) rather than Assumption~(P1).
-
-\paragraph{Proof of Theorem~6.}
-Follows identically to the proof of Theorem~1, substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~7 rather than Proposition~3, Proposition~8 rather than Proposition~4, and Assumption~(P6) rather than Assumption~(P1).
-
-\subsection{Testing Results}
-
-\paragraph{Proof of Theorem~2.}
-We have already upper bounded the Type I error of $\varphi$ in Lemma~\ref{lem:fixed_graph_testing}, and it remains to upper bound the Type II error.
-To do so, we condition on the event that the design points $X_1,\ldots,X_n$ satisfy,
-\begin{equation}
-\label{pf:laplacian_eigenmaps_testing_fo_1}
-\dotp{L_{n,\varepsilon}f_0}{f_0}_n \leq \frac{C}{\delta}M^2,\quad \textrm{and} \quad \lambda_k \geq \min\{\lambda_k(\Delta_P), \varepsilon^{-2}\}~~\textrm{for all $2 \leq k \leq n$,}
-\end{equation}
-as well as that
-\begin{equation}
-\label{pf:laplacian_eigenmaps_testing_fo_2}
-\|f_0\|_n^2 \geq \frac{1}{2}\|f_0\|_P^2.
-\end{equation}
-Note that by Propositions~3 and~4, both statements in~\eqref{pf:laplacian_eigenmaps_testing_fo_1} are satisfied with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\}$.
-Additionally, by Proposition~6 and the assumption that $\|f_0\|_P^2 \geq CM^2/(bn^{2/d})$, the one-sided inequality~\eqref{pf:laplacian_eigenmaps_testing_fo_2} follows with probability at least $1 - \exp\{-(cn \wedge 1/b)\}$.
-Setting $\delta = b/3$ and taking $n \geq N$ to be sufficiently large, the bottom line is that both~\eqref{pf:laplacian_eigenmaps_testing_fo_1} and~\eqref{pf:laplacian_eigenmaps_testing_fo_2} are together satisfied with probability at least $1 - b/2$.
-
-Now, to complete the proof of Theorem~2, we would like to invoke Lemma~\ref{lem:fixed_graph_testing}, and conclude that conditional on $X_1,\ldots,X_n$ satisfying~\eqref{pf:laplacian_eigenmaps_testing_fo_1} and~\eqref{pf:laplacian_eigenmaps_testing_fo_2}, our test $\varphi$ will equal $1$ with probability at least $1 - b/2$.
-To use Lemma~\ref{lem:fixed_graph_testing}, we will need to establish that~\eqref{eqn:fixed_graph_testing_critical_radius} is satisfied, which we now show.
-
-On the one hand, we have that the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius} is upper bounded, 
-\begin{align*}
-\frac{\dotp{L_{n,\varepsilon}f_0}{f_0}_n}{\lambda_{K + 1}} + \frac{\sqrt{2K}}{n}\biggl[2\sqrt{\frac{1}{a}} + \sqrt{\frac{2}{b}} + \frac{32}{bn}\biggr] & \leq C\biggl(\frac{M^2}{b \min\{\lambda_{K+1}(\Delta_P), \varepsilon^{-2}\}} + \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr) \\
-& \leq C\biggl(\frac{M^2}{b}K^{-2/d} + \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr)
-\end{align*}
-with the second inequality following by the assumption $\varepsilon \leq K^{-1/d}$ and Weyl's Law.
-On the other hand, we have that $\|f_0\|_n^2 \geq \|f_0\|_P^2/2$.
-Consequently, to prove Theorem~2, it remains only to verify that
-\begin{equation}
-\label{pf:laplacian_eigenmaps_testing_fo_3}
-\|f_0\|_P^2 \geq C\biggl(\frac{M^2}{b}K^{-2/d} + \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr).
-\end{equation}
-As in the estimation case, we can further upper bound the right hand side of~\eqref{pf:laplacian_eigenmaps_testing_fo_3}, depending on the value of $K$ chosen in~(P2).
-The classical case is $K = (M^2n)^{d/(2 + d)}$, in which case~\eqref{pf:laplacian_eigenmaps_testing_fo_3} is satisfied as long as
-\begin{equation*}
-\|f_0\|_P^2 \geq CM^2(M^2n)^{-4/(4 + d)}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]
-\end{equation*}
-If $M^2 < n^{-1}$, then we take $K = 1$, and~\eqref{pf:laplacian_eigenmaps_testing_fo_3} is satisfied whenever
-\begin{equation*}
-\|f_0\|_P^2 \geq \frac{C}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr].
-\end{equation*}
-Finally if $M > n^{1/d}$, we take $K = n$, and~\eqref{pf:laplacian_eigenmaps_testing_fo_3} is satisfied if
-\begin{equation*}
-\|f_0\|_P^2 \geq C\biggl(\frac{M^2}{n^{2/d}b} + n^{-1/2}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr).
-\end{equation*}
-We conclude by observing that by assumption, $\|f_0\|_P^2$ implies each of these three inequalities, and thus implies~\eqref{pf:laplacian_eigenmaps_testing_fo_3}.
-
-\paragraph{Proof of Theorem~4.}
-Follows identically to the proof of Theorem~1, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~4 rather than Proposition~3 and Assumption~(P4) rather than Assumption~(P2).
-
-\paragraph{Proof of Theorem~7.}
-Follows identically to the proof of Theorem~1, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~7 rather than Proposition~3, Proposition~8 rather than Proposition~4, Proposition~9 rather than Proposition~6, and Assumption~(P6) rather than Assumption~(P2).
-
-\paragraph{Proof of Theorem~5.}
-Note that our choices of $K$ and $\varepsilon$ ensure  that~\eqref{pf:laplacian_eigenmaps_testing_fo_1} (with $L_{n,\varepsilon}^s$ replacing $L_{n,\varepsilon}$) and~\eqref{pf:laplacian_eigenmaps_testing_fo_2} are satisfied with probability at least $1 - b/2$.
-Proceeding as in the proof of Theorem~2, we upper bound the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius},
-\begin{align*}
-\frac{\dotp{L_{n,\varepsilon}f_0}{f_0}_n}{\lambda_{K + 1}} + \frac{\sqrt{2K}}{n}\biggl[2\sqrt{\frac{1}{a}} + \sqrt{\frac{2}{b}} + \frac{32}{bn}\biggr] & \leq C\biggl(\frac{M^2}{b \min\{\lambda_{K+1}(\Delta_P), \varepsilon^{-2}\}} + \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr) \\
-& \leq C\biggl(\frac{M^2}{b}\varepsilon^2 + \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr).
-\end{align*}
-Unlike in the proof of Theorem~2, we note that in this case $\varepsilon^2 \leq C\lambda_K(\Delta_P)$ rather than vice versa.
-From here, proceeding as in the proof of Theorem~2 gives the claimed result.
-
-\section{Analysis of kernel smoothing}
+\section{Proofs}
+\section{Proofs}
 \label{subsec:kernel_smoothing}
-In this section we prove Lemma~1 (in Section~\ref{subsec:pf_kernel_smoothing_insample}) and Lemma~\ref{lem:kernel_smoothing_bias} (in Section~\ref{subsec:pf_kernel_smoothing_bias}).
+In this section we prove Lemma~\ref{lem:kernel_smoothing_insample} (in Section~\ref{subsec:pf_kernel_smoothing_insample}) and Lemma~\ref{lem:kernel_smoothing_bias} (in Section~\ref{subsec:pf_kernel_smoothing_bias}).
 In Section~\ref{subsec:eigenmaps_beats_kernel_smoothing}, we give a sequence of design densities and regression functions $\{p^{(n)}(x), f_0^{(n)}(x)\}_{n \in \mathbb{N}}$ for which the estimator $T_{n,h}\wc{f}$ (extension of Laplacian eigenmaps by kernel smoothing) strictly outperforms directly kernel smoothing the responses, in the sense that
@@
 This expression reflects the known fact that the bias operator of kernel smoothing is equal to the non-local Laplacian, up to a rescaling by the population degree functional $d_{P,h}(x)$.
-In the second equality, we are using the notation $L_{P,h}f(x)$ exactly as defined in equation~(35) of the main text, but with the kernel $\psi$ instead of $\eta$.
+In the second equality, we are using the notation $L_{P,h}f(x)$ exactly as defined in~\eqref{eqn:nonlocal_laplacian}, but with the kernel $\psi$ instead of $\eta$.
 Note that $\psi$ satisfies all the same assumptions as $\eta$, except that of positivity; when $\psi$ is a higher-order kernel it may take negative values.
@@
 \end{align*}
-Since by assumption $\mc{X}$ has Lipschitz boundary, setting $c_0$ to be a sufficiently small constant in~(P7), we can further deduce that $\int \psi(\|z\|) \1\{hz + x \in \mc{X}\} \,dz \geq 1/3$, and consequently that
+Since by assumption $\mc{X}$ has Lipschitz boundary, setting $c_0$ to be a sufficiently small constant in~\ref{asmp:bandwidth}, we can further deduce that $\int \psi(\|z\|) \1\{hz + x \in \mc{X}\} \,dz \geq 1/3$, and consequently that
 \begin{equation}
@@
 
-\subsection{Proof of Lemma~1}
+\subsection{Proof of Lemma~\ref{lem:kernel_smoothing_insample}}
 \label{subsec:pf_kernel_smoothing_insample}
@@
 	\label{lem:kernel_smoothing_bias}
-	Suppose Model~1, and that the kernel smoothing operator $T_{P,h}$ is computed with respect to a kernel $\eta$ that satisfies~(K3).
+	Suppose Model~\ref{def:model_flat_euclidean}, and that the kernel smoothing operator $T_{P,h}$ is computed with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel}.
 	\begin{itemize}
@@
 		\end{equation*}
-		\item If $f_0 \in H_0^{s}(\mc{X})$, $p \in C^{s - 1}(\mc{X})$, and $\eta$ satisfies~(K4), then there exists a constant $C$ which does not depend on $f_0$ such that
+		\item If $f_0 \in H_0^{s}(\mc{X})$, $p \in C^{s - 1}(\mc{X})$, and $\eta$ satisfies~\ref{asmp:ho_kernel}, then there exists a constant $C$ which does not depend on $f_0$ such that
 		\begin{equation*}
@@
 
-\subsection{Comparison of Laplacian Eigenmaps to kernel smoothing}
-\label{subsec:eigenmaps_beats_kernel_smoothing}
-
-In this section, we give a sequence of densities $p^{(n)}$ and regression functions $f_0^{(n)}$ for which the $L^2(P)$ error of the two-stage estimator $T_{n,h}\wh{f}$ is stochastically dominated by the $L^2(P)$ error of applying kernel smoothing directly to the responses.
-We begin by precisely setting up the problem and reviewing the two estimators we will compare.
-Then we give and prove our formal claim.
-
-\paragraph{Setup.}
-As usual, we assume the design points $(X_1,\ldots,X_n)$ are sampled independently from a distribution $P$, and the responses
-\begin{equation*}
-Y_i = f_0(X_i) + \varepsilon_i,
-\end{equation*}
-where the noise terms $\varepsilon_i \sim N(0,1)$ are sampled independently, and are also independent of the design points.
-
-Unlike in our main text, we shall assume a specific (parametric) form for $p = p^{(n)}$ and $f_0 = f_0^{(n)}$, that changes as a function of $n$. More specifically, using the notation
-\begin{equation*}
-r := \frac{(\log n)^2}{n}, \quad Q_1 := [0,1/2 - r], \quad Q_2 := [1/2 + r,1],
-\end{equation*}
-we take the domain $\mc{X}^{(n)} = Q_1 \cup Q_2$, and 
-\begin{equation}
-\label{def:model_cluster_assumption}
-p^{(n)}(x) := \frac{1}{1 - 2r}\1\bigl\{x \in Q_1 \cup Q_2\bigr\}, \quad f_0^{(n)}(x) := \theta \cdot \Bigl(\1\bigl\{x \in Q_1\bigr\} - \1\bigl\{x \in Q_2\bigr\}\Bigr),
-\end{equation}
-where $\theta \in \Reals$.
-Note that as $n \to \infty$, the density $p^{(n)}$ is converging (in, say, $L^2([0,1])$ norm) to the uniform distribution on $[0,1]$, and the regression function $f_0^{(n)}$ to a step function.
-
-Our goal is to estimate $f_0$ using an estimator $f$, in such a way that the resulting $L^2(P)$ error $\|f - f_0\|_P^2$ is small.
-We shall consider two estimators: Laplacian eigenmaps + extension by kernel smoothing, and direct kernel smoothing of the responses.
-In the notation of our main text, these are respectively $T_{n,h}\wh{f}$ and $T_{n,h'}{\bf Y}$.
-For simplicity, in both cases we shall consider only the boxcar kernel,
-\begin{equation}
-\label{asmp:boxcar_kernel}
-\eta(z) = \psi(z) = \1\{z \leq 1\}.
-\end{equation}
-
-\paragraph{Laplacian eigenmaps + kernel smoothing beats direct kernel smoothing.}
-Now we are in a position to state our result. Let $P^{(n)}$ be the distribution with density $p^{(n)}$, i.e. the uniform probability distribution on $Q_1 \cup Q_2$. 
-\begin{proposition}
-	\label{prop:eigenmaps_beats_kernel_smoothing}
-	Suppose $(X_1,Y_1),\ldots(X_n,Y_n)$ are sampled according to~\eqref{def:model_cluster_assumption}. 
-	\begin{itemize}
-		\item Compute the Laplacian eigenmaps estimator $\wh{f}$ using a kernel $\eta$ which satisfies~\eqref{asmp:boxcar_kernel}, number of eigenvectors $K = 2$, and radius $\varepsilon = ((\log n)^2/n)/2$.
-Compute the kernel smoothed extension $T_{n,h}\wh{f}$ using a kernel $\psi$ which satisfies~\eqref{asmp:boxcar_kernel}, and bandwidth $h = ((\log n)^2/n)/2$.
-Then,
-		\begin{equation}
-		\label{eqn:eigenmaps_beats_kernel_smoothing_1}
-		\Ebb \|T_{n,h}\wh{f} - f_0^{(n)}\|_{L^2(P^{(n)})}^2 \leq \frac{8 (\log n)^2 + 6 (\log n)^2 \theta^2}{n}.
-		\end{equation}
-		\item Compute the kernel smoothing estimator $T_{n,h}{\bf Y}$, using a kernel $\psi$ which satisfies~\eqref{asmp:boxcar_kernel}. Then
-		\begin{equation}
-		\label{eqn:eigenmaps_beats_kernel_smoothing_2}
-		\inf_{h' > 0} \Ebb\Bigl[\|T_{n,h'}{\bf Y} - f_0^{(n)}\|_{L^2(P^{(n)})}^2\Bigr] \geq \min\biggl\{ \frac{1}{16(\log n)^2}, \frac{\theta}{\sqrt{32 n}} \biggr\}.
-		\end{equation}
-	\end{itemize}
-\end{proposition}
-The statements~\eqref{eqn:eigenmaps_beats_kernel_smoothing_1} and~\eqref{eqn:eigenmaps_beats_kernel_smoothing_2} are finite sample and hold for any $\theta$.
-Taking a sequence $(\theta_n)_{n \in \mathbb{N}}$ such that $\theta = o(\sqrt{n})$ and $\theta = \omega((\log n)^2/\sqrt{n})$ implies that
-\begin{equation*}
-\lim_{n \to \infty} \sup_{h'}  \frac{\Ebb \|T_{n,h}\wh{f} - f_0^{(n)}\|_{P^{(n)}}^2}{	\Ebb \|T_{n,h'}{\bf Y} - f_0^{(n)}\|_{P^{(n)}}^2 } = 0.
-\end{equation*}
-
-\paragraph{Proof of Proposition~\ref{prop:eigenmaps_beats_kernel_smoothing}.}
-First we prove~\eqref{eqn:eigenmaps_beats_kernel_smoothing_1}, then~\eqref{eqn:eigenmaps_beats_kernel_smoothing_2}.
-
-\underline{\emph{Proof of~\eqref{eqn:eigenmaps_beats_kernel_smoothing_1}}.}
-We begin by showing that with high probability the eigenvectors $v_1,v_2$ respect the cluster structure of $p$.
-Formally, letting $u_1 = (\1\{X_i \in Q_1\})_{i \in [n]}$, and likewise $u_2 = (\1\{X_i \in Q_2\})_{i \in [n]}$, we establish that
-\begin{equation}
-\label{pf:eigenmaps_beats_kernel_smoothing_1}
-\Pbb\Bigl(\mathrm{span}\{v_1,v_2\} = \mathrm{span}\{u_1,u_2\}\Bigr) \geq 1 - \frac{1}{n}.
-\end{equation}
-Note that~\eqref{pf:eigenmaps_beats_kernel_smoothing_1} is equivalent to the statement that the neighborhood graph $G_{n,\varepsilon}$ consists of exactly two connected components: one consisting of all design points $X_i \in Q_1$, and the other consisting of all design points $X_i \in Q_2$.
-Of course, since $h < r$ and $\eta$ is compactly supported on $[0,1]$, it follows that no $X_i \in Q_1$ and $X_j \in Q_2$ can be connected.
-On the other hand, using an elementary concentration argument (stated in Lemma~\ref{lem:balls_in_bins}) and the triangle inequality, we deduce that with probability at least $1 - 1/n$ there exists a path in $G_{n,\varepsilon}$ between each $X_i, X_j \in Q_1$, and likewise between each $X_i,X_j \in Q_2$.
-This establishes~\eqref{pf:eigenmaps_beats_kernel_smoothing_1}.
-
-Now we condition on the ``good'' event $\mc{E}$ that the design points $X_1,\ldots,X_n$ satisfy~\eqref{eqn:balls_in_bins}, and therefore that $\mathrm{span}\{v_1,v_2\} = \mathrm{span}\{u_1,u_2\}$.
-Consider the empirical mean $\wb{Y}_Q := \frac{1}{\sharp\{Q \cup {\bf X}\}} \sum_{i: X_i \in Q} Y_i$.
-Since $\mathrm{span}\{v_1,v_2\} = \mathrm{span}\{u_1,u_2\}$, the estimator $\wh{f}$ will be piecewise constant on $Q_1$ and $Q_2$, and in fact we have that
-\begin{equation*}
-\wh{f} = \wb{Y}_{Q_1}u_1 + \wb{Y}_{Q_2}u_2.
-\end{equation*}
-Moreover, by~\eqref{eqn:balls_in_bins} every $x \in Q_1$ is within distance $h$ of at least one $X_i \in {\bf X} \cup Q_1$ and likewise every $x \in Q_2$ is within distance $h$ of at least one $X_j \in {\bf X} \cup Q_2$.
-Consequently, the kernel smoothening of $\wh{f}$ also exhibits a piecewise constant structure,
-\begin{equation*}
-T_{n,h}\wh{f}(x) = \wb{Y}_{Q_1}\1\{x \in Q_1\} + \wb{Y}_{Q_2}\1\{x \in Q_2\},\quad \textrm{for all}~~x \in Q_1 \cup Q_2.
-\end{equation*}
-Letting
-\begin{equation*}
-\|T_{n,h}\wh{f} - f_0^{(n)}\|_{L^2(P^{(n)})}^2 = \frac{1}{2}\Bigl((\wb{Y}_{Q_1} - \theta)^2 + (\wb{Y}_{Q_2} + \theta)^2\Bigr),
-\end{equation*}
-it follows from Lemma~\ref{lem:chi_square_bound} that conditional on $\mc{E}$,
-\begin{equation}
-\label{pf:eigenmaps_beats_kernel_smoothing_0}
-\|T_{n,h}\wh{f} - f_0^{(n)}\|_{L^2(P^{(n)})}^2 \leq \frac{4(\log n)^2}{n}.
-\end{equation}
-
-Now we derive a crude upper bound on $\|T_{n,h}\wh{f} - f_0^{(n)}\|_{L^2(P^{(n)})}$ that will suffice to control the error conditional on $\mc{E}^c$. Note that if $d_{n,h}(x) = 0$ then
-$T_{n,h}\wh{f}(x) = 0$, and $\bigl(T_{n,h}\wh{f}(x) - f_0(x)\bigr)^2 = [f_0(x)]^2$. On the other hand, if $d_{n,h}(x) \geq 1$, then by Jensen's inequality it follows that
-\begin{align*}
-\bigl(T_{n,h}\wh{f}(x) - f_0(x)\bigr)^2 & \leq \biggl(\frac{1}{d_{n,h}(x)} \sum_{i = 1}^{n} (\wh{f}_i - f_0(x)) \cdot \1\{\|X_i - x\| \leq h \}\biggr)^2 \\
-& \leq \frac{1}{d_{n,h}(x)} \sum_{i = 1}^{n} (\wh{f}_i - f_0(x))^2 \cdot \1\{\|X_i - x\| \leq h \} \\
-& \leq 2[f_0(x)]^2 + \frac{2}{d_{n,h}(x)} \sum_{i = 1}^{n} (\wh{f}_i)^2 \cdot \1\{\|X_i - x\| \leq h\},
-\end{align*}
-and integrating over all $x \in \mc{X}^{(n)}$ gives
-\begin{align*}
-\|T_{n,h}\wh{f} - f_0\|_{L^2(P^{(n)})}^2 & \leq 2 \|f_0\|_{L^2(P^{(n)})}^2 + \sum_{i = 1}^{n} (\wh{f}_i)^2 \biggl(\int \frac{\1\{\|X_i - x\| \leq h \}}{d_{n,h}(x)} \,dP(x)\biggr) \\
-& \leq 2 \|f_0\|_{L^2(P^{(n)})}^2 + 2h\sum_{i = 1}^{n} (\wh{f}_i)^2 \\
-& = 2 \|f_0\|_{L^2(P^{(n)})}^2 + 2 (\log n)^2 \|\wh{f}\|_n^2.
-\end{align*}
-We can further upper bound the empirical norm of $\|\wh{f}\|_n^2$ using the Cauchy-Schwarz inequality:
-\begin{equation*}
-\|\wh{f}\|_n^2 \leq \frac{2}{n}\sum_{i = 1}^{n}\dotp{Y}{v_1}_n^2v_{1,i}^2 + \dotp{{\bf Y}}{v_2}_n^2v_{2,i}^2 \leq 2\bigl(\dotp{{\bf Y}}{v_1}_n^2 + \dotp{{\bf Y}}{v_2}_n^2\bigr) \leq 4 \|{\bf Y}\|_n^2.
-\end{equation*}
-Noting that $\Ebb[\|{\bf Y}\|_n^2|{\bf X}] = \|f_0\|_n^2 + 1/n$, and that $\max_x |f_0(x)| = \theta$, we conclude that
-\begin{equation*}
-\Ebb\Bigl[\|T_{n,h}\wh{f} - f_0\|_{L^2(P^{(n)})}^2 \cdot \1\{\mc{E}^c\}\Bigr] \leq \Ebb\Bigl[\Bigl(2 \|f_0\|_{L^2(P^{(n)})}^2 + 4 (\log n)^2 \|{\bf Y}\|_n^2\Bigr) \cdot \1\{\mc{E}^c\} \Bigr] \leq \frac{2 \theta^2}{n} + \frac{4 (\log n)^2(\theta^2 + 1)}{n}.
-\end{equation*}
-Combining this with~\eqref{pf:eigenmaps_beats_kernel_smoothing_0} implies~\eqref{eqn:eigenmaps_beats_kernel_smoothing_1}.
-
-\underline{\emph{Proof of~\eqref{eqn:eigenmaps_beats_kernel_smoothing_2}}.}
-Fix $x \in \mc{X}^{(n)}$. A standard argument using the law of iterated expectation implies the following lower bound on the pointwise risk in terms of squared-bias and variance-like quantities,
-\begin{equation*}
-\Ebb\Bigl[\Bigl(T_{n,h'}{\bf Y}(x) - f_0(x)\Bigr)^2\Bigr] \geq \Ebb\biggl[\Bigl(f_0(X) - f_0(x)\Bigr)^2|X \in B(x,h')\biggr] + \Ebb\biggl[\frac{1}{d_{n,h'}(x)}\biggr].
-\end{equation*}
-The variance term can be lower bounded quite simply for any $x \in \mc{X}^{(n)}$; noting that $\sup_{x} p^{(n)}(x) < 2$ and $\nu(B(x,h') \cap \mc{X}^{(n)}) \leq 2h'$, it follows by Jensen's inequality that
-\begin{equation*}
-\Ebb\biggl[\frac{1}{d_{n,h'}(x)}\biggr] \geq \frac{1}{\Ebb[d_{n,h'}(x)]} \geq \frac{1}{4nh'}.
-\end{equation*}
-On the other hand the squared bias term is quite large for $x$ close to $1/2$. Precisely, if $h' \geq 4r$ then a simple calculation implies
-\begin{equation*}
-\Ebb[(f_0(X) - f_0(x))^2|X \in B(x,h')] \geq \frac{\theta^2}{8} \quad \textrm{for all}~x \in [(1 - h'/2)_{+}, 1/2 - r].
-\end{equation*}
-Combining these lower bounds on variance and squared bias terms and integrating over $x' \in \mc{X}$, we deduce the following: if $h' \leq 4r$, then
-\begin{equation*}
-\Ebb\Bigl[\|T_{n,h'}{\bf Y} - f_0^{(n)}\|_{L^2(P^{(n)})}^2\Bigr] \geq \frac{1}{16(\log n)^2},
-\end{equation*}
-whereas if $h' > 4r$  then
-\begin{equation*}\Ebb\Bigl[\|T_{n,h'}{\bf Y} - f_0^{(n)}\|_{L^2(P^{(n)})}^2\Bigr] \geq \frac{1}{4nh'} + \frac{\theta^2h'}{32}.
-\end{equation*}
-In the latter case, setting the derivative equal to $0$ shows that the right hand side is always at least $\theta/\sqrt{32 n}$, and taking the minimum over the two cases then yields~\eqref{eqn:eigenmaps_beats_kernel_smoothing_2}.
-
-\paragraph{A Useful Lemma.}
-Let $m = 4n/\log^2n$. Furthermore, for $i = 0,\ldots,m - 1$, let
-\begin{equation*}
-Q_{i1} = [i/m,(i + 1)/m] \cdot (1/2 - r), \quad Q_{i2} = 1/2 + [i/m,(i + 1)/m] \cdot (1/2 - r).
-\end{equation*}
-\begin{lemma}
-	\label{lem:balls_in_bins}
-	Suppose $(X_1,Y_1),\ldots(X_n,Y_n)$ are sampled according to~\eqref{def:model_cluster_assumption}. For any $\log(n) > 16$, we have that 
-	\begin{equation}
-	\label{eqn:balls_in_bins}
-	\Pbb\Bigl(\sharp\{Q_{ij} \cup {\bf X}\} > 0 ~~\textrm{for all $i = 1,\ldots,m - 1$ and $j = 1,2$} \Bigr) \geq 1 - \frac{1}{n}.
-	\end{equation}
-\end{lemma}
-Let $\varepsilon = h = \log^2n/2n$.
-Note that by construction,~\eqref{eqn:balls_in_bins} implies that any points $x$ and $x'$ in adjacent intervals $Q_{ij}$ and $Q_{i'j}$ must be connected in $G_{n,\varepsilon}$.
-It also implies that $d_{n,h}(x) > 0$ for every $x \in Q_1 \cup Q_2$.
-
-\paragraph{Proof (of Lemma~\ref{lem:balls_in_bins}).}
-For each $Q_{ij}$, we have that 
-\begin{equation*}
-\Pbb\bigl(\sharp\{Q_{ij} \cup {\bf X}\} = 0\bigr) = (1 - 1/(2m))^{n} \leq \exp\{-n/2m\}.
-\end{equation*}
-Thus by a union bound,
-\begin{equation*}
-\Pbb\Bigl(\sharp\{Q_{ij} \cup {\bf X}\} = 0 ~~\textrm{for any $i = 1,\ldots,m - 1$ and $j = 1,2$} \Bigr) \leq 2m\exp\{-n/2m\} = \frac{8n}{\log^2n}\exp\{-\log^2n/8\},
-\end{equation*}
-and the claim follows from some basic algebra.
-
-\section{Miscellaneous}
-Here we give assorted helpful Lemmas used at various points in the above proofs. We also review notation and relevant facts regarding Taylor expansion.
-
-\subsection{Concentration Inequalities}
-Lemma~\ref{lem:chi_square_bound} controls the deviation of a chi-squared random variable. It is from~\cite{laurent00}.
-\begin{lemma}
-	\label{lem:chi_square_bound}
-	Let $\xi_1,\ldots,\xi_N$ be independent $N(0,1)$ random variables, and let $U := \sum_{k = 1}^{N} a_k(\xi_k^2 - 1)$.  Then for any $t > 0$,
-	\begin{equation*}
-	\Pbb\Bigl[U \geq 2 \|a\|_2 \sqrt{t} + 2 \|a\|_{\infty}t\Bigr] \leq \exp(-t).
-	\end{equation*}
-	In particular if $a_k = 1$ for each $k = 1,\ldots,N$, then
-	\begin{equation*}
-	\Pbb\Bigl[U\geq 2\sqrt{N t} + 2t\Bigr] \leq \exp(-t).
-	\end{equation*}
-\end{lemma}
-
-Lemma~\ref{lem:one_sided_bernstein} is an immediate consequence of the one-sided Bernstein's inequality (14.23) in \cite{wainwright2019}.
-\begin{lemma}[One-sided Bernstein's inequality]
-	\label{lem:one_sided_bernstein}
-	Let $X, X_1,\ldots,X_n \sim P$, and $f$ satisfy $\Ebb[f^4(X)] < \infty$. Then
-	\begin{equation*}
-	\|f\|_n^2 \geq \frac{1}{2}\|f\|_P^2,
-	\end{equation*}
-	with probability at least $1 - \exp\bigl(-n/8 \cdot \|f\|_P^4 /\Ebb[f^4(X)]\bigr)$.
-\end{lemma}
-
-In the proof of Lemma~1, we require uniform control of the empirical degree functional $d_{n,h}(x)$ over all $x \in \mc{X}$.
-Such a result is available to us because the kernel $\psi$ is Lipschitz on its support, so that the class of functions $\{\psi((x - \cdot)/h): h \in \Rd\}$ has finite VC dimension.
-The precise estimate we use is due to~\cite{gine2002}.
-\begin{lemma}[Uniform bound for empirical degree.]
-	\label{lem:uniform_bound_empirical_degree}
-	Suppose Model~1. For a kernel $\psi$ satisfying~(K3) and bandwidth $h$ satisfying~(P7), there exist constants $c, C, c_1$ and $C_1$ which do not depend on $h$ or $n$ such that
-	\begin{equation*}
-	\Pbb\biggl(\sup_{x \in \Rd} n \cdot \Bigl|d_{n,h}(x) - d_{P,h}(x)\Bigr| > t\biggr) \leq C\exp\biggl(-c \frac{t^2}{nh^d}\biggr),
-	\end{equation*}
-	for any $t \in \Reals$ satisfying
-	\begin{equation}
-	\label{eqn:uniform_bound_empirical_degree_1}
-	C_1\sqrt{nh^d \log(1/h)} \leq t \leq c_1 nh^d.
-	\end{equation}
-\end{lemma}
-Now we translate Lemma~\ref{lem:uniform_bound_empirical_degree} into a multiplicative bound, which will be more useful for our purposes. 
-Recall the lower bound on $d_{P,h}(x)$ given in~\eqref{eqn:degree_lower_bound}.
-By setting $C_0$ to be a sufficiently large constant in~(P7), we can ensure that choosing $t = n h^d p_{\min}/6$ satisfies both the inequalities~\eqref{eqn:uniform_bound_empirical_degree_1}.
-For this choice of $t$ it follows from~\eqref{eqn:degree_lower_bound} and Lemma~\ref{lem:uniform_bound_empirical_degree} that
-\begin{equation}
-\label{eqn:uniform_bound_empirical_degree_2}
-\sup_{x \in \Rd} \biggl|\frac{d_{n,h}(x) - d_{P,h}(x)}{d_{P,h}(x)}\biggr| \leq \sup_{x \in \Rd} \biggl|\frac{d_{n,h}(x) - d_{P,h}(x)}{2t/n}\biggr| \leq \frac{1}{2}
-\end{equation}
-with probability at least $1 - C\exp(-cnh^d)$. This is the form of the result we use in the proof of Lemma~1.
-
-\subsection{Taylor expansion}
-\label{subsec:taylor_expansion}
-We begin with some notation that allows us to concisely derivatives.
-For a given $z \in \Rd$ and $s$-times differentiable function $f: \mc{X} \to \Reals$, we denote $\bigl(d_x^sf\bigr)(z) := \sum_{|\alpha| = s} D^{\alpha}f(x) z^{\alpha}$.
-We also write $d^sf := \sum_{|\alpha| = j} D^{\alpha}f$.
-We point out that in the first-order case $d_x^1f$ is the differential of $f$ at $x \in \mc{X}$, while $d^1f$ is the divergence of $f$.
-
-Let $u$ be a function which is $s$ times continuously differentiable at all $x \in \mc{X}$, for $k \in \mathbb{N}\setminus\{0\}$.
-Suppose that for some $h > 0$, $x \in \mc{X}_{h}$ and $x' \in B(x,h)$.
-We write the order-$s$ Taylor expansion of $u(x')$ around $x' = x$ as
-\begin{equation*}
-u(x') = u(x) + \sum_{j = 1}^{s - 1} \frac{1}{j!}\bigl(d_x^{j}u\bigr)(x' - x) + r_{x'}^{s}(x;u)
-\end{equation*}
-For notational convenience we have adopted the convention that $\sum_{j = 1}^{0} a_j = 0$.
-Thus $\bigl(d_x^{j}f\bigr)(z)$ is a degree-$j$ polynomial---and so a $j$-homogeneous function---in $z$, meaning for any $t \in \Reals$,
-\begin{equation*}
-\bigl(d_x^{j}f\bigr)(tz) = t^{j} \cdot \bigl(d_x^{j}f\bigr)(z).
-\end{equation*}
-The remainder term $r_{x'}$ is given by
-\begin{equation*}
-r_{x'}^s(x;f) = \frac{1}{(j - 1)!} \int_{0}^{1}(1 - t)^{j - 1} \bigl(d_{x + t(x' - x)}^{s}f\bigr)(x' - x) \,dt,
-\end{equation*}
-where we point out that the integral makes sense because $x + t(x' - x) \in B(x,h) \subseteq \mc{X}$.
-We now give estimates on the remainder term in both sup-norm and $L^2(\mc{X}_{h})$ norm, each of which hold for any $z \in B(0,1)$.
-In sup-norm, we have that
-\begin{equation*}
-\sup_{x \in \mc{X}_{h}}|r_{x + hz}^j(x;f)| \leq C h^j \|f\|_{C^j(\mc{X})},
-\end{equation*}
-whereas in $L^2(\mc{X}_{h})$ norm we have,
-\begin{equation}
-\label{eqn:sobolev_remainder_term}
-\int_{\mc{X}_{h}} \bigl|r_{x + thz}^j(x;f)\bigr|^2 \,dx \leq h^{2j} \int_{\mc{X}_{h}} \int_{0}^{1} |d_{x + thz}^jf(z)|^2 \,dt \,dx \leq h^{2j} \|d^jf\|_{L^2(\mc{X})}^2.
-\end{equation}
-In the last inequality
-
-Finally, we recall some facts regarding the interaction between smoothing kernels and polynomials.
-Let $q_j(z)$ be an arbitrary degree-$j$ (multivariate) polynomial.
-If $\eta$ is a radially symmetric kernel and $j$ is odd, then by symmetry it follows that
-\begin{equation*}
-\int_{B(0,1)} q_j(z) \eta(\|z\|) \,dz = 0.
-\end{equation*}
-On the other hand, if $\psi$ is an order-$s$ kernel for some $s > j$, then by converting to polar coordinates we can verify that
-\begin{equation*}
-\int_{B(0,1)} q_j(z) \eta(\|z\|) \,dz = 0.
-\end{equation*}
-
-\section{Computational considerations}
-\label{sec:computational_considerations}
-Recall that when $s = 1$, we have shown that Laplacian eigenmaps is optimal when $\varepsilon \asymp (\log n/n)^{1/d}$ is (up to a constant) as small as possible while still ensuring the graph $G$ is connected.
-On the other hand, when $s > 1$, we can show Laplacian eigenmaps is optimal only when $\varepsilon = \omega(n^{-c})$ for some $c < 1/d$.
-For such a choice of $\varepsilon$, the average degree in $G$ will grow polynomially in $n$ as $n \to \infty$, and computing eigenvectors of the Laplacian of a graph will be more computationally intensive than if the graph were sparse.
-In this dense-graph setting, we now discuss a procedure to more efficiently compute an approximation to the Laplacian eigenmaps estimate: \emph{edge sparsification}.
-
-By now there exist various methods see (e.g., the seminal papers of \citet{spielman2011,spielman2013,spielman2014}, or the overview by \citet{vishnoi2012} and references therein) to efficiently remove many edges from the graph $G$ while only slightly perturbing the spectrum of the Laplacian.
-Specifically such algorithms take as input a parameter $\sigma \geq 1$, and return a sparser graph $\wc{G}$, $E(\wc{G}) \subseteq E(G)$, with a Laplacian $\wc{L}_{n,\varepsilon}$ satisfying
-\begin{equation*}
-\frac{1}{\sigma} \cdot u^{\top} \wc{L}_{n,\varepsilon} u \leq u^{\top} L_{n,\varepsilon} u \leq \sigma \cdot u^{\top} \wc{L}_{n,\varepsilon}u \quad \textrm{for all $u \in \Reals^n$.}
-\end{equation*}
-Let $\wc{f}$ be the Laplacian eigenmaps estimator computed using the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ .
-Because $\wc{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wc{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$, and consequently much faster to compute $\wc{f}$ than $\wh{f}$.
-Statistically speaking, letting $\wc{\lambda}_k$ be the $k$th eigenvalue of $\wc{L}_{n,\varepsilon}$, we have that conditional on $\{X_1,\ldots,X_n\}$,
-\begin{equation*}
-\|\wc{f} - f_0\|_n^2 \leq \frac{\dotp{\wc{L}_{n,\varepsilon}^s f_0}{f_0}_n}{\wc{\lambda}_{K + 1}^s} + \frac{5K}{n} \leq \sigma^{2s} \frac{\dotp{\wc{L}_{n,\varepsilon}^s f_0}{f_0}_n}{\wc{\lambda}_{K + 1}^s} + \frac{5K}{n},
-\end{equation*}
-with probability at least $1 - \exp(-K)$.
-Consequently $\wt{f}$ has $L^2(P_n)$-error of at most $\sigma^{2s}$ times our upper bound on the error of $\wh{f}$, and for any choice of $\sigma$ that is constant in $n$ the estimator $\wc{f}$ will also be rate-optimal.
-
-In fact the aforementioned edge sparsification algorithms are overkill for our needs.
-For one thing, they are designed to work when $\sigma$ is very close to $1$, whereas in order for $\wc{f}$ to be rate-optimal, setting $\sigma$ to be any constant greater than $1$, say $\sigma = 2$, is sufficient.
-Additionally, edge sparsification algorithms are traditionally designed to work in the worst-case, where no assumptions are made on the structure of the graph $G$.
-But the geometric graphs we consider in this paper exhibit a special structure, in which very roughly speaking no single edge is a bottleneck.
-As pointed out by~\citet{sadhanala16b}, in this special case there are far simpler and faster methods for sparsification, which at least empirically seem to do the job.
-
-% Bibliography
-\bibliographystyle{imsart-nameyear}
-\bibliography{../../../graph_regression_bibliography}
-
 \end{document}

### 2021-12-22 notes/technical_notes/estimation_graph_sobolev.tex
Some miscellaneous notes.

\documentclass{article}

\input{preamble.sty}

\begin{document}
	% Front Matter
	\title{Estimation over Graph Sobolev Spaces}
	\author{Alden Green}
	\maketitle

	\tableofcontents

	\section{Introduction}
	We consider a Normal means problem. We observe ${\bf Y} = (Y_1,\ldots,Y_n) \in \Reals^n$ according to the signal plus Gaussian noise model,
	\begin{equation}
	Y_i = \theta_i^{\ast} + w_i,
	\end{equation}
	where $w_i \sim N(0,1)$ are independent Gaussian noise variables. The task is to recover the mean vector $\theta_i^{\ast} \in \Reals^n$, which is generally impossible unless some structure is assumed.

	The classical approach assumes that $\theta_i^{\ast}$ exhibits \emph{decay} as a function of $i$. We take a different route, and assume \emph{smoothness} of the mean vector $\theta^{\ast}$ with respect to some graph $\texttt{G} = (\texttt{V},\texttt{E})$. This smoothness is defined with respect to the edge incidence matrix $D \in \Reals^{m \times n}$---which has rows $D_{e} = (0,\ldots,1,\ldots,-1,\ldots,0)$ for each edge $e = (i,j) \in \ttt{E}$, and $m = |E|$ is the number of edges in $\ttt{G}$---and the Laplacian $L = D^{\top}D$. In particular, we assume $\theta^{\ast}$ belongs to a \emph{graph Sobolev class}, which we now define.

	\begin{definition}[Graph Sobolev Space]
		For a given $s \in \mathbb{N}, s \geq 1$ and $p \in [1,\infty)$, the \emph{graph Sobolev seminorm} of a vector $\theta \in \Reals^n$ is given by
		\begin{equation}
		|\theta|_{s,p} := 
		\begin{cases}
		\|DL^{(s - 1)/2}\|_p, & \quad\textrm{$s$ odd.}\\
		\|L^{s/2}\|_p, & \quad\textrm{$s$ even.}
		\end{cases}
		\end{equation}
		The graph Sobolev class $W_n^{s,p} \subseteq \Reals^n$ consists of all vectors $\theta \in \Reals^n$ for which $\|\theta\|_{s,p} < \infty$. The graph Sobolev ball $W_{n}^{s,p}(M) = \{\theta \in W_n^{s,p}: |\theta|_{s,p} < M\}$.
	\end{definition} 
	We address the following questions:
	\begin{itemize}
		\item \emph{Minimax rates}. The risk of an estimator $\wh{\theta} \in \Reals^n$--i.e. a measurable function of ${\bf Y}$---with respect to mean-squared error loss is
		\begin{equation}
		\label{eqn:risk}
		R\bigl(\wh{\theta},\theta^{\ast}\bigr) = \frac{1}{n}\Ebb[\|\wh{\theta} - \theta^{\ast}\|_2^2]. 
		\end{equation}
		The minimax risk over balls $W_{n}^{s,p}(M)$ is
		\begin{equation}
		R(W_{n}^{s,p}(M)) = \inf_{\wh{\theta}} \sup_{\theta^{\ast} \in W_n^{s,p}(M)} 	R\bigl(\wh{\theta},\theta^{\ast}\bigr).
		\end{equation}
		We would like to find a function $\psi(n,M)$ such that $R(W_{n}^{s,p}(M)) \asymp \psi(n,M)$.
		\item \emph{Adaptivity}. \textcolor{red}{(TODO)}.
		\item \emph{Fractional Sobolev spaces} \textcolor{red}{(TODO)}.
	\end{itemize}
	Our particular angle will be to get as far as we can using only discrete analogues to classical results of Fourier analysis.

	\section{Discrete Fourier Analysis}
	The Laplacian $L$ is a positive semi-definite matrix, and according to the spectral theorem can be characterized as $L = \sum_{k = 1}^{n} \lambda_k v_k v_k^{\top}$, where the eigenvalue/eigenvector pairs $(\lambda_k,v_k)$ satisfy
	\begin{equation}
	L v_k = \lambda_k v_k, \quad \|v_k\|_2^2 = 1,
	\end{equation} 
	and are index in ascending order of eigenvalues $0 = \lambda_1 \leq \lambda_2 \leq \ldots \leq \lambda_n$. We assume the spectrum of $L$ exhibits some structure characteristic of certain geometric graphs.
	\begin{assumption}{A}{1}
		\label{asmp:weyl}
		There exist positive numbers $a$ and $A$ such that
		\begin{equation}
		\label{eqn:weyl}
		a k^{2/d} \leq \lambda_k \leq A K^{2/d}, \textrm{for all $k \in \{2,\ldots,n\}$.}
		\end{equation}
	\end{assumption}
	\begin{assumption}{A}{2}
		\label{asmp:laplacian-spectrum} 
		The graph Laplacian $L$ of $\ttt{G}$ satisfies the following: there exists a positive number $B$ such that
		\begin{equation}
		\label{eqn:eigenvector-incoherence}
		\max_{k = 1,\ldots,n} \|v_k\|_{\infty} \leq \frac{B}{\sqrt{n}}.
		\end{equation}
	\end{assumption}
	Trivially~\eqref{eqn:weyl} is satisfied for any connected graph $\ttt{G}$, and~\eqref{eqn:eigenvector-incoherence} for any graph, connected or not. However, we will obtain sharp rates for estimation only when $a \asymp A$ and $B \asymp 1$.

	Let $F: \Reals^n \to \Reals^n$ denote the Graph Fourier Transform $F\theta := (\theta^{\top} v_1,\ldots,\theta^{\top}v_n)$.  Under Assumption~\ref{asmp:laplacian-spectrum}, we can reproduce many results of classical Fourier analysis.
	\begin{proposition}
		For any graph $\ttt{G}$ with incidence matrix $D$, Laplacian $L$, and GFT $F$, the following statements hold for all $\theta \in \Reals^n$.
		\begin{itemize}
			\item \emph{Parseval}. $\|\theta\|_2 = \|F\theta\|_2$
			\item \emph{Laplacian to Multiplication}. $F(L\theta) = \vec{\lambda} \odot F(\theta)$. Here $\vec{\lambda} := (\lambda_1,\ldots,\lambda_n)$ and for vectors $v,u \in \Reals^n, v \odot u = (v_1u_1,\ldots,v_nu_n)$ stands for pointwise multiplication.
		\end{itemize}
		Suppose additionally that $\ttt{G}$ satisfies~\ref{asmp:laplacian-spectrum}. Then
		\begin{itemize}
			\item \emph{Duality}. $\|\theta\|_1 \leq (B/\sqrt{n}) \|F\theta\|_{\infty}$.
			\item \emph{Hausdorff-Young}.  
		\end{itemize}
	\end{proposition}
\end{document}

### 2021-12-22 talks/laplacian_smoothing/MURI/laplacian_smoothing.tex
Added the talk I gave to the MURI regarding Laplacian smoothing.

\documentclass[xcolor=dvipsnames]{beamer}

% Packages
\usepackage{bm}
\usepackage{wasysym}

%% Theme related.
\usetheme{default}
\setbeamerfont{author}{size = \large}
\setbeamerfont{institute}{size=\large}
\setbeamerfont{footline}{series = \bfseries}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{frametitle}[default][center]
\setbeamertemplate{itemize items}[circle]
\definecolor{Red}{rgb}{.9,.05,.05}
\setbeamercolor{alerted text}{fg=Red}
\usefonttheme[onlymath]{serif}

%% Math
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\wt}[1]{\widetilde{#1}}
\DeclareMathOperator*{\argmin}{argmin}

%% Font
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}

\title{Nonparametric Regression via Graph Laplacian Smoothing}

\author{Alden Green}
\institute{Department of Statistics and Data Science \\
\includegraphics[width=.5 \textwidth]{figures/cmu.png}}
\date{}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\addtobeamertemplate{navigation symbols}{}{
	\usebeamerfont{footline}
	\usebeamercolor[fg]{footline}
	\hspace{1em}
	\insertframenumber
}
\addtocounter{framenumber}{-1}

%%% (1) Introduction
% (A) Method

\begin{frame}{Graph Laplacian smoothing}
We observe: 
\begin{itemize}
	\item Design points $X_1,\ldots,X_n \in \Omega$ (sampled from distribution $P$ supported on $\Omega \subseteq \Reals^D$).
	\item Responses $\vec{Y} = (Y_1,\ldots,Y_n)$.
\end{itemize}
\begin{figure}
	\includegraphics[width=.6\textwidth, trim = {0 3cm 0 1.5cm}]{figures/warmup_example_1.pdf}
\end{figure}
\pause
{\bf Idea}: Solve a penalized least squares problem with penalty involving Laplacian of a neighborhood graph built over the design.
\end{frame}

\begin{frame}[t]{Graph Laplacian smoothing}
Form a {\bf neighborhood graph} $G = (\{X_1,\ldots,X_n\}, W)$. 
\begin{itemize}
	\item Compute adjacency matrix $W$ using kernel $\eta$ and radius $\varepsilon$, so
	\begin{equation*}
	W_{ij} = \eta\biggl(\frac{\|X_i - X_j\|}{\varepsilon}\biggr)
	\end{equation*}
	measures similarity between points $X_i$ and $X_j$. 
	\item E.g. weighted $\varepsilon$-geometric graph: $W_{ij} = 1(\|X_i - X_j\| \leq \varepsilon) \cdot \phi(\|X_i - X_j\|/\varepsilon)$.
\end{itemize}
\pause
\begin{figure}
	\centering
	\includegraphics[width=.6\textwidth, trim = {0 3cm 0 1.5cm}]{figures/warmup_example_2.pdf}
\end{figure}
\end{frame}

\begin{frame}[t]{Graph Laplacian smoothing}
Compute the {\bf Laplacian matrix} $L = D - W$, where 
\begin{equation*}
D_{ii} = \sum_{j = 1}^{n} W_{ij},
\end{equation*}
is the degree of $X_i$ in $G$.
\vspace{.2cm}
\begin{figure}
	\centering
	\includegraphics[height=.4\textheight]{figures/laplacian_matrix.png}
\end{figure}
\end{frame}

\begin{frame}[t]{Graph Laplacian smoothing}
Solve the penalized least squares problem
\begin{equation*}
\wh{\theta} := \argmin_{\theta \in \Reals^n} \frac{1}{n}\sum_{i = 1}^{n}(Y_i - \theta_i)^2 + \lambda \theta^{\top} L \theta.
\end{equation*}
{\bf Intuition}: The penalty $\theta^{\top} L \theta = \sum_{i,j = 1}^{n} (\theta_i - \theta_j)^2 W_{ij}$ enforces a notion of smoothness.
\pause
\begin{figure}
	\centering
	\includegraphics[width=.6\textwidth, trim = {0 3cm 0 1.5cm}]{figures/warmup_example_3.pdf}
\end{figure}
\end{frame}

% (C) Motivation
\begin{frame}{Laplacian smoothing: motivation}
Why Laplacian smoothing?
\begin{itemize}
	\item \emph{Computational ease.} Sparse graph, fast solvers.
	\item \emph{Generality.} Makes sense whenever one can form a graph $G$. 
	\item \emph{Weak supervision.} Naturally extends to semi- and unsupervised problems.
\end{itemize}
\pause
Today I will focus on...
\begin{itemize}
	\item \emph{Univariate intuition}. In $1d$, discretization of spline-based estimator.
	\item \emph{Multivariate extension}. Depends in a sensible manner on geometric structure of design.
	\item \emph{Statistically optimal}. Achieves minimax rates over Sobolev classes.
\end{itemize}
\end{frame}

\begin{frame}[t]{Laplacian smoothing: motivation}
In $1d$, recall the first-order~\alert{smoothing spline} estimator
\begin{equation*}
\widetilde{f} := \argmin_{f \in H^1} \|Y - f\|_n^2 + \lambda \int (f')^2.
\end{equation*}
\pause
As $n \to \infty, \varepsilon \to 0$, for smooth functions $f$ (say $f \in C^{1,\alpha}$), 
\begin{equation*}
f^{\top} L f = \frac{1}{2}\sum_{i,j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 W_{ij} \approx \int \bigl(f')^2 p^2 .
\end{equation*}
\pause
{\bf Interpretation}: Laplacian smoothing serves as a \alert{discretized} approximation to a~\alert{density-weighted} smoothing spline. 
\begin{itemize}
	\item Density-weighting useful under (hard or soft) cluster assumption, or manifold hypothesis.
	\item Discretization useful when density $p$ is unknown, and for multivariate extension.
\end{itemize}
\end{frame}

% (B) Setup
\begin{frame}{Basic problem setup}

{\bf Nonparametric regression, random design}: We observe $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n \in \Reals^D$ are sampled independently from density $p$, and
\begin{equation*}
Y_i = f_0(X_i) + Z_i, \quad Z_i \sim N(0,1).
\end{equation*}
Goal is to accurately estimate the unknown regression function $f_0$, which is assumed to be first-order \alert{Sobolev} smooth, with semi-norm
\begin{equation*}
|f_0|_{H^1}^2 := \int \|\nabla f_0(x)\|^2 \,dx \leq 1.
\end{equation*}
{\bf Loss}: Measured in \alert{empirical norm}, meaning
\begin{equation*}
\|\wh{\theta} - f_0\|_n^2 = \frac{1}{n}\sum_{i = 1}^{n} \bigl(\wh{\theta}_i - f_0(X_i)\bigr)^2.
\end{equation*}
\end{frame}

% (2) Results

\begin{frame}[t]{Rates of convergence: univariate case}

{\bf Background}: First-order smoothing splines are known to achieve minimax optimal rates of convergence over first-order Sobolev classes.
\begin{equation*}
\sup_{f_0 \in H^1(\Omega)} \mathbb{E}\|\wt{f} - f_0\|_{L^2(\Omega)}^2 \asymp n^{-2/3} \asymp \inf_{\wh{f}}\sup_{f_0 \in H^1(\Omega)} \mathbb{E}\|\wt{f} - f_0\|_{L^2(\Omega)}^2
\end{equation*}

{\bf Question}: Can Laplacian smoothing match this rate of convergence?

\end{frame}

\begin{frame}[t]{Rates of convergence: univariate case}

{\bf Result}: Yes. In fact, Laplacian smoothing matches these rates so long as $\Omega$ has \alert{intrinsic dimension} $d = 1$ (no matter the ambient dimension).
\vspace{.2 in}
\pause

Main idea behind result: 
\begin{itemize}
	\item \alert{Bias-variance} trade off involving eigenvalues $\rho_k(L)$,
	\begin{equation*}
	\|\wh{\theta} - f_0\|_n^2 \lesssim \lambda |f_0|_{H^1}^2 + \sum_{k = 1}^{n}\frac{1}{(1 + \lambda \rho_k(L))^2}.
	\end{equation*}
	\item Under conditions on $\Omega$ and $P$, eigenvalues satisfy \alert{modified Weyl's Law} scaling, $\rho_k(L) \asymp k^{2} \wedge \varepsilon^{-2}$.
	\item Choosing $\varepsilon \asymp (\log n/n)$ and $\lambda \asymp n^{-2/3}$ balances bias and variance.
\end{itemize}
\pause 
{\bf Interpretation}: Laplacian smoothing is a statistically reasonable way to adapt a classic variational scheme (first-order smoothing splines) to complex geometric structures. 
\end{frame}

\begin{frame}[t]{Rates of convergence: multivariate case}

{\bf Background}: Extension of first-order smoothing splines to multivariate setting---replacing derivative by gradient,
\begin{equation*}
\min_{f_0 \in H_1} \|\vec{Y} - f\|_n^2 + \lambda \int \|\nabla f\|^2 \,dx,
\end{equation*}
does not lead to a well-posed method.\pause
\begin{itemize}
	\item \alert{Sobolev embedding theorem}: Convergence in Sobolev norm $\not\Longrightarrow$ pointwise convergence. There exist functions $f_k$ satisfying $f_k(X_i) = Y_i$ yet 
	$$\lim_{k \to \infty} \int \|\nabla f_k\|_2^2 = 0.$$
\end{itemize}
\pause
{\bf Question}: Any hope for Laplacian smoothing?
\end{frame}

\begin{frame}[t]{Rates of convergence: multivariate case}

{\bf Result}: Yes. 
\begin{itemize}
	\item Laplacian smoothing achieves optimal rates of convergence ($n^{-2/(2 + d)}$) so long as $\Omega$ has intrinsic dimension $2 \leq d \leq 4$.
	\item Rate of convergence is $n^{-4/(3d)}$ when $d \geq 5$. Suboptimal.
\end{itemize}
\pause
Main idea behind result: 
\begin{itemize}
	\item Bias-variance tradeoff involving eigenvalues $\rho_k(L)$,
	\begin{equation*}
	\|\wh{\theta} - f_0\|_n^2 \lesssim \lambda |f_0|_{H^1}^2 + \sum_{k = 1}^{n}\frac{1}{(1 + \lambda \rho_k(L))^2}.
	\end{equation*}
	\item Under conditions on $\Omega$ and $P$, eigenvalues satisfy modified Weyl's Law scaling, $\rho_k(L) \asymp k^{2/d} \wedge \varepsilon^{-2}$.
	\item Choosing $\varepsilon \asymp (\log n/n)^{1/d}$ and $\lambda \asymp n^{-2/(2+ d)}$ balances bias and variance when $2 \leq d \leq 4$.
\end{itemize}
\pause
{\bf Interpretation}: ???
\end{frame}

\begin{frame}
\centering
\huge
\blue{Questions}
\end{frame}

% (3) Questions
\begin{frame}[t]{Questions}
\begin{itemize}
	\item Can we situate Laplacian smoothing within the standard regression toolbox?
	\item Is there a conceptually coherent way to extend the estimator to be defined over all of $\Omega$?
\end{itemize}
\textbf{My thought}: Both of these questions are (at least somewhat) related to properties of interpolation.
\end{frame}

% (3A): Equivalent kernel
\begin{frame}[t]{Equivalent kernel of Laplacian smoothing}

\textbf{Question 1}: Can we situate Laplacian smoothing within the standard regression toolbox?
\pause

\textbf{Goal}: Recast Laplacian smoothing as approximately the estimate produced by smoothing with a particular convolutional kernel. \\
\vspace{.2 in}

\begin{itemize}
	\item Mathematically, this means a kernel function $\kappa(\cdot)$ and a bandwidth $h(\cdot)$ such that
	\begin{equation*}
	\wh{\theta}_i \approx \frac{1}{n} \sum_{j = 1}^{n} \kappa\biggl(\frac{\|X_i - X_j\|}{h(X_i)}\biggr) Y_j, \quad~\textrm{for all $i = 1,\ldots,n$.}
	\end{equation*}
\end{itemize}
\pause
Laplacian smoothing matrix $S_i$ satisfies the first-order equation
\begin{equation*}
(I + \lambda L)S_i = e_i.
\end{equation*}
What does $S_i$ look like?

\end{frame}

% Labeled
\begin{frame}[t]{Equivalent kernel of Laplacian smoothing}
\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/equivalent_kernel/laplacian_smoothingequivalent_kernel_uniform.pdf}
	\includegraphics[width=.49\textwidth]{figures/equivalent_kernel/laplacian_smoothingequivalent_kernel_gaussian_mixture.pdf}
\end{figure}
\pause
\textbf{Takeaways}
\begin{itemize}
	\item \alert{Impulse} at point $X_i$, assigning much more weight to response $Y_i$ than to any other $Y_j,~j \neq i$.
\end{itemize}
Why?
\end{frame}

\begin{frame}[t]{Equivalent kernel of Laplacian smoothing: Details}
\textbf{Impulse}. Laplacian smoothing equivalent kernel $S_i$ satisfies the first-order equation
\begin{equation*}
(I + \lambda L)S_i = e_i.
\end{equation*}
\pause
Some simple algebra yields $S_i = \wt{e}_i + \wt{S}_i$, where
\begin{itemize}
	\item \alert{Impulse} $\wt{e}_i$ is given by
	\begin{equation*}
	\wt{e}_i = (I + \lambda D)^{-1} e_i = \frac{1}{1 + \lambda D_{ii}} e_i.
	\end{equation*}
	\item Smooth \alert{filter} $\wt{S}_i$ satisfies a modified first-order equation,
	\begin{equation*}
	(I + \lambda L)\wt{S}_i = \lambda W (I + \lambda D)^{-1} e_i = \frac{\lambda}{(1 + \lambda D_{ii})} W_i.
	\end{equation*}
\end{itemize}
What does $\wt{S}_i$ look like?
\end{frame}

% Unlabeled
\begin{frame}[t]{Equivalent kernel of Laplacian smoothing}
\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/equivalent_kernel/unlabeled_laplacian_smoothingequivalent_kernel_uniform.pdf}
	\includegraphics[width=.49\textwidth]{figures/equivalent_kernel/unlabeled_laplacian_smoothingequivalent_kernel_gaussian_mixture.pdf}
\end{figure}
\textbf{Takeaways}
\begin{itemize}
	\item \alert{Bandwidth} is density-dependent. Smaller in low-density areas, larger in high-density areas.
	\item \alert{Shape} approximates Laplace kernel (?)
\end{itemize}
\end{frame}

\begin{frame}[t]{Equivalent kernel of Laplacian smoothing: Details}
\textbf{Bandwidth, intuition}. In $1d$, the Laplacian smoothing optimization problem is a discretization of 
\begin{equation*}
\min_{f \in H^1}\int (y - f)^2 \textcolor{red}{p} + \lambda \int (f')^2 \textcolor{red}{p^2}.
\end{equation*}
\begin{itemize}
	\item When $p(x)$ is large, places more weight on smoothness (large bandwidth).
	\item When $p(x)$ is small, places more weight on data-fitting (small bandwidth).
\end{itemize}
\pause
Note: recall that the first-order smoothing spline $\wt{f}$ approximately solves
\begin{equation*}
\min_{f \in H^1}\int (y - f)^2 \textcolor{red}{p} + \lambda \int (f')^2.
\end{equation*}
$\Longrightarrow$ Opposite dependence of bandwidth on density. 
\end{frame}

\begin{frame}[t]{Equivalent kernel of Laplacian smoothing: Details}
\textbf{Shape, intuition}. The equivalent kernel $\wt{S}_i$ satisfies a discretization of the following differential equation:
\begin{equation}
\label{eqn:diffeq}
(I + \lambda \Delta_P) \wt{s}_i = \eta(\|X_i - \cdot\|/\varepsilon),
\end{equation}
where $\Delta_P$ is a density-weighted Laplacian operator
\begin{equation*}
\Delta_Pf = -\frac{1}{p} \mathrm{div}(p^2 \nabla f).
\end{equation*}
\pause
\begin{itemize}
	\item In $1d$, when $p$ is uniform, fundamental solution to~\eqref{eqn:diffeq} is a Laplace kernel. Equivalent kernel of first-order smoothing spline.
	\item In $d \geq 2$, when $p$ is uniform, fundamental solution to~\eqref{eqn:diffeq} is a singular kernel. 
\end{itemize}
\end{frame}

\begin{frame}[t]{Equivalent kernel of Laplacian smoothing: Details}
\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/equivalent_kernel/laplacian_smoothing_grid2d.pdf}
	\includegraphics[width=.49\textwidth]{figures/equivalent_kernel/unlabeled_laplacian_smoothing_grid2d.pdf}
\end{figure}
\end{frame}

% (3B): Extension
% Words: Technique for finding numerical approximation to eigenfunction problems.
\begin{frame}[t]{Laplacian smoothing: Nystr\"{o}m method}
\textbf{Question 2}: Is there a \alert{sensible} way to extend $\wh{\theta} \in \Reals^n$ to $\Omega$?
\pause

\textbf{Background}: Nystr\"{o}m method. Technique for approximately solving eigenfunction problems
\begin{equation*}
\int \eta(x,y) \phi(y) \,dy = \rho \phi(x),
\end{equation*}
by discretization. \pause
\begin{itemize}
	\item Let $W_{ij} = \eta(X_i,X_j)$. Solve the system of equations 
	$$\frac{1}{n}\sum_{j = 1}^{n} W_{ij} \phi_j = \rho \phi_i, \quad i = 1,\ldots,n,$$
	to produce estimate $(\hat{\phi}_1,\ldots,\hat{\phi}_n)$.
	\item Plug back in to get the~\alert{Nystr\"{o}m extension},
	\begin{equation*}
	\wh{\phi}(x) = \frac{1}{\rho n}\sum_{j = 1}^{n} \eta(x,x_j)\wh{\phi}_j.
	\end{equation*}
\end{itemize}
\end{frame}

% armed with our understanding...
\begin{frame}[t]{Laplacian smoothing: Nystr\"{o}m method}

{\bf Impulse + filter}: Some algebra shows Laplacian smoothing estimate $\wh{\theta} = D_{\lambda}^{-1}\vec{Y} + \wt{\theta}$, where $D_{\lambda} = (I + \lambda D)$, and $\wt{\theta}$ satisfies
\begin{equation*}
\begin{aligned}
(I + \lambda L) \wt{\theta} & = \lambda W D_{\lambda}^{-1}\vec{Y}.
\end{aligned}
\end{equation*}
{\bf Idea}: Apply Nystr\"{o}m method to $\wt{\theta}$.
\begin{itemize}
	\item Solve the system of equations
	\begin{equation*}
	(I + \lambda D_{ii})\theta_i = \lambda \sum_{j = 1}^{n} W_{ij}\Bigl(\frac{Y_j}{1 + \lambda D_{jj}} + \theta_j\Bigr), \quad i = 1,\ldots,n.
	\end{equation*}
	to produce estimate $(\wt{\theta}_1,\ldots,\wt{\theta}_n)$.
	\item Plug back in to get the \alert{Nystr\"{o}m extension},
	\begin{equation*}
	\wt{\theta}(x) = \frac{\lambda}{1 + \lambda D(x)}\sum_{j = 1}^{n} \eta\biggl(\frac{\|x - X_j\|}{\varepsilon}\biggr) \biggl(\frac{Y_j}{1 + \lambda D_{jj}} + \wt{\theta}_j\biggr).
	\end{equation*}
\end{itemize}
\end{frame}

\begin{frame}[t]{Laplacian smoothing: Nystr\"{o}m method}
\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/warmup_example_3.pdf}
	\includegraphics[width=.49\textwidth]{figures/warmup_example_4.pdf}
\end{figure}
\end{frame}

\begin{frame}
\centering
\Huge Thank you!
\end{frame}

\end{document}
